\chapter{Hamiltonian Evolution and Time-Sliced Path Integrals}
\label{ch:hamiltonian-evolution-time-sliced-path-integrals}

\section{Classical Lagrangian and Hamiltonian Data}

Let \(Q\) be a smooth \(d\)-dimensional configuration manifold.  Local
coordinates on \(Q\) are denoted by \(q^a\), \(a=1,\ldots,d\).  A tangent
vector is written \((q,\dot q)\in TQ\), and a cotangent vector is written
\((q,p)\in T^*Q\), with \(p=p_a\,\dd q^a\).  A classical Lagrangian is a
function
\[
  L:TQ\to\mathbb R.
\]
For a path \(q:[t_i,t_f]\to Q\), the action is
\[
  S[q]=\int_{t_i}^{t_f}L(q(t),\dot q(t))\,\dd t .
\]
The variational principle with fixed endpoints gives the Euler--Lagrange
equations
\[
  \frac{\dd}{\dd t}\frac{\partial L}{\partial \dot q^a}
  =
  \frac{\partial L}{\partial q^a}.
\]
Thus a Lagrangian description uses the pair \((q,\dot q)\) as local Cauchy
data for the classical second-order equation.

Assume now that the fiber derivative of \(L\),
\[
  (q,\dot q)\longmapsto
  \left(q,p_a=\frac{\partial L}{\partial\dot q^a}\right),
\]
is invertible on the region under consideration.  The Hamiltonian associated
to this regular Lagrangian is the Legendre transform
\[
  H(q,p)=p_a\dot q^a-L(q,\dot q)
  \quad
  \text{with }\dot q=\dot q(q,p).
\]
The canonical Poisson bracket on \(T^*Q\) is
\[
  \{f,g\}_{\rm P}
  =
  \frac{\partial f}{\partial q^a}\frac{\partial g}{\partial p_a}
  -
  \frac{\partial f}{\partial p_a}\frac{\partial g}{\partial q^a},
\]
where repeated indices are summed.  Hamiltonian time evolution is
\[
  \dot f=\{f,H\}_{\rm P},
\]
and, in particular,
\[
  \dot q^a=\frac{\partial H}{\partial p_a},
  \qquad
  \dot p_a=-\frac{\partial H}{\partial q^a}.
\]
The Lagrangian and Hamiltonian descriptions are equivalent only after the
regularity hypothesis above has been imposed; without it, constraints must be
included as part of the classical data.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[>=Stealth,node distance=1.7cm,
  every node/.style={font=\small}]
  \node[draw,rounded corners=2pt,align=center,inner sep=4pt] (lagdata)
    {Lagrangian data\\ \((q,\dot q)\)};
  \node[draw,rounded corners=2pt,align=center,inner sep=4pt,right=of lagdata]
    (action) {Action\\ \(S=\int L(q,\dot q)\,\dd t\)};
  \node[draw,rounded corners=2pt,align=center,inner sep=4pt,right=of action]
    (eom) {Equations\\ \(\delta S=0\)};
  \node[draw,rounded corners=2pt,align=center,inner sep=4pt,below=of action]
    (hamdata) {Hamiltonian data\\ \((q,p)\), \(H(q,p)\)};
  \draw[->] (lagdata) -- (action);
  \draw[->] (action) -- (eom);
  \draw[<->] (action) -- node[right,font=\scriptsize]
    {\(p_a=\partial L/\partial\dot q^a\)} (hamdata);
  \node[align=left,font=\scriptsize,right=0.55cm of hamdata] {
    \(H=p_a\dot q^a-L\)\\[2pt]
    \(\dot q^a=\partial H/\partial p_a\)\\[2pt]
    \(\dot p_a=-\partial H/\partial q^a\)};
\end{tikzpicture}
\caption{Classical Lagrangian and Hamiltonian descriptions, related by a
regular Legendre transform.}
\label{fig:classical-lagrangian-hamiltonian-data}
\end{figure}

\section{Canonical Quantum-Mechanical Data}

Let \(d\) be the number of degrees of freedom and let
\(\mathcal Q=\mathbb R^d\) be the configuration space. The Hilbert space is
\[
  \Hilb_{\mathcal Q}=L^2(\mathcal Q,\dd^dq).
\]
The coordinate operators \(\widehat q^a\), \(a=1,\ldots,d\), act by
multiplication. The momentum operators are
\[
  \widehat p_a=-\ii\hbar\,\frac{\partial}{\partial q^a}
\]
first on the common invariant test domain
\(\mathcal S(\mathbb R^d)\subset\Hilb_{\mathcal Q}\), and satisfy
\[
  [\widehat q^a,\widehat p_b]=\ii\hbar\,\delta^a_b.
\]
Closures or self-adjoint extensions are specified when a Hamiltonian is
chosen.  The Schwartz-domain convention is enough for the distributional kernel
manipulations below; the operator limit is then stated at the Hilbert-space
level.

Let \(\widehat H\) be a self-adjoint Hamiltonian on \(\Hilb_{\mathcal Q}\).
The time-evolution operator over time \(T\in\mathbb R\) is
\[
  U(T)=\exp\left(-\frac{\ii}{\hbar}T\widehat H\right).
\]
In the Schr\"odinger picture, a vector evolves as
\[
  \ket\psi\longmapsto U(t)\ket\psi .
\]
In the Heisenberg picture, an operator \(\widehat O\) evolves as
\[
  \widehat O(t)=U(t)^{-1}\widehat O\,U(t),
  \qquad
  \frac{\dd}{\dd t}\widehat O(t)
  =
  \frac{\ii}{\hbar}[\widehat H,\widehat O(t)]
\]
on vectors for which the commutator expression is defined.  The path integral
constructed below gives a regulated representation of matrix elements of this
Hamiltonian time evolution.

The position and momentum generalized eigenstates are denoted by
\(\ket{q}\) and \(\ket{p}\). They are elements of
\(\mathcal S'(\mathbb R^d)\) in the rigged Hilbert space
\[
  \mathcal S(\mathbb R^d)
  \subset
  L^2(\mathbb R^d,\dd^dq)
  \subset
  \mathcal S'(\mathbb R^d),
\]
with the \(\hbar\)-dependent Fourier convention
\[
  \widetilde\psi(p)
  =
  \int_{\mathbb R^d}\dd^dq\,
  \ee^{-\frac{\ii}{\hbar}p_aq^a}\psi(q).
\]
They are normalized by
\[
  \braket{q}{p}=\ee^{\frac{\ii}{\hbar}p_aq^a},
\]
and used through the resolutions
\[
  \int_{\mathbb R^d}\dd^dq\,\ket{q}\bra{q}=1,
  \qquad
  \int_{\mathbb R^d}\frac{\dd^dp}{(2\pi\hbar)^d}\,\ket{p}\bra{p}=1.
\]
These are weak identities: after inserting test vectors on the left and right
they become the ordinary Fourier inversion formula.
The transition amplitude is the distributional kernel
\[
  K(q_f,T;q_i,0)
  =
  \bra{q_f}U(T)\ket{q_i}.
\]
The continuum Lagrangian path-integral notation for this same kernel is
\[
  \bra{q_f}U(T)\ket{q_i}
  =
  \int_{q(0)=q_i}^{q(T)=q_f} [Dq]\,
  \exp\left(\frac{\ii}{\hbar}S[q]\right).
\]
This displayed equation is a compact name for the regulated construction
below.  The symbol \([Dq]\) is determined by the regulator, limiting procedure,
operator symbol, and boundary conditions; at fixed regulator the object is a
finite-dimensional oscillatory integral.

\section{Finite Time Slicing}

Choose an integer \(N\ge1\), set \(\epsilon=T/N\), and define
\[
  t_n=n\epsilon,\qquad n=0,\ldots,N,
\]
with \(q_0=q_i\) and \(q_N=q_f\). Inserting \(N-1\) position resolutions and
\[
  U(T)=\left[\exp\left(-\frac{\ii}{\hbar}\epsilon\widehat H\right)\right]^N
\]
gives first the position-space finite product
\[
  K(q_f,T;q_i,0)
  =
  \int\prod_{n=1}^{N-1}\dd^dq_n\,
  \prod_{n=0}^{N-1}
  \bra{q_{n+1}}
  \exp\left(-\frac{\ii}{\hbar}\epsilon\widehat H\right)
  \ket{q_n}.
\]
Inserting \(N\) momentum resolutions gives
\[
\begin{aligned}
  K(q_f,T;q_i,0)
  &=
  \int\prod_{n=1}^{N-1}\dd^dq_n
  \prod_{n=0}^{N-1}
  \frac{\dd^dp_n}{(2\pi\hbar)^d}
  \\
  &\quad\times
  \prod_{n=0}^{N-1}
  \bra{q_{n+1}}p_n\rangle
  \bra{p_n}\exp\left(-\frac{\ii}{\hbar}\epsilon\widehat H\right)\ket{q_n}.
\end{aligned}
\]

\begin{center}
\begin{tikzpicture}[>=Stealth]
  \draw[line width=0.6pt] (0,0) -- (9.2,0);
  \foreach \x/\lab in {
    0/{q_0=q_i},
    1.25/{q_1},
    2.5/{q_2},
    3.75/{q_3},
    6.75/{q_{N-1}},
    8.75/{q_N=q_f}
  }{
    \draw (\x,0.11) -- (\x,-0.11);
    \node[above] at (\x,0.11) {$\lab$};
  }
  \node[below] at (0,-0.12) {$t_0=0$};
  \node[below] at (8.75,-0.12) {$t_N=T$};
  \node at (5.2,0.19) {$\cdots$};
  \draw[decorate,decoration={brace,amplitude=4pt,mirror}]
    (0,-0.5) -- node[below=3pt] {$\epsilon=T/N$} (1.25,-0.5);
\end{tikzpicture}
\end{center}

Suppose a symbol \(h(q,p)\) for \(\widehat H\) has been fixed by an operator
ordering convention, for example by
\[
  \bra{p}\widehat H\ket{q}=h(q,p)\braket{p}{q}
\]
on the chosen domain.  Replacing each short-time matrix element by the first
terms of its symbol expansion gives
\[
  \bra{q_{n+1}}p_n\rangle
  \bra{p_n}\ee^{-\ii\epsilon\widehat H/\hbar}\ket{q_n}
  =
  \exp\left[
    \frac{\ii}{\hbar}
    \left(
      p_{n,a}(q_{n+1}^a-q_n^a)
      -\epsilon h(q_n,p_n)
      +O(\epsilon^2)+O(\hbar\epsilon)
    \right)
  \right]
\]
for this convention, with the error understood in the regulated
operator-symbol expansion rather than as a pointwise statement about an
undefined continuum integral.  The associated finite-\(N\) phase-space
approximant is
\[
  K_N(q_f,T;q_i,0)
  =
  \int
  \prod_{n=1}^{N-1}\dd^dq_n
  \prod_{n=0}^{N-1}\frac{\dd^dp_n}{(2\pi\hbar)^d}
  \exp\left(\frac{\ii}{\hbar}S_N[p,q]\right),
\]
where the discrete phase-space action is
\[
  S_N[p,q]
  =
  \sum_{n=0}^{N-1}
  \left[
    p_{n,a}(q_{n+1}^a-q_n^a)
    -\epsilon\,h(q_n,p_n)
  \right].
\]
The continuum notation
\[
  \int Dq\,Dp\,
  \exp\left[
    \frac{\ii}{\hbar}\int_0^T
    \bigl(p_a\dot q^a-h(q,p)\bigr)\dd t
  \right]
\]
denotes this regulated construction together with the specified limiting
procedure. In the present chapter, the finite-\(N\) expression is the defined
object.

\section{Trotter Limits and Kernel Convergence}

The passage from the finite product to a continuum path integral is a
convergence statement about operators before it is a statement about kernels.
Let
\[
  \widehat H=\widehat A+\widehat B
\]
be a self-adjoint Hamiltonian obtained from two self-adjoint pieces on a
specified common form domain.  Suppose the hypotheses of the Trotter product
formula hold for this sum.  Then, for real time,
\[
  \exp\left(-\frac{\ii T}{\hbar}\widehat H\right)
  =
  \operatorname*{s-lim}_{N\to\infty}
  \left[
    \exp\left(-\frac{\ii\epsilon}{\hbar}\widehat A\right)
    \exp\left(-\frac{\ii\epsilon}{\hbar}\widehat B\right)
  \right]^N,
  \qquad
  \epsilon=\frac{T}{N},
\]
where the limit is in the strong operator topology on \(\Hilb_{\mathcal Q}\).
The Euclidean version is the semigroup limit
\[
  \exp\left(-\frac{\tau}{\hbar}\widehat H\right)
  =
  \operatorname*{s-lim}_{N\to\infty}
  \left[
    \exp\left(-\frac{\delta}{\hbar}\widehat A\right)
    \exp\left(-\frac{\delta}{\hbar}\widehat B\right)
  \right]^N,
  \qquad
  \delta=\frac{\tau}{N},
\]
when the corresponding semibounded form construction applies.

For the standard Hamiltonian
\[
  \widehat H=\frac{1}{2m}\widehat p^2+V(\widehat q),
\]
the product formula says that the operator obtained by alternating free
propagation and multiplication by the potential converges to the exact
evolution operator.  Inserting position resolutions between the factors gives
a finite-dimensional integral at each \(N\).  Thus the time-sliced path
integral is a representation of a product formula.

The distributional kernel
\[
  \bra{q_f}\exp\left(-\frac{\ii T}{\hbar}\widehat H\right)\ket{q_i}
\]
requires an additional topology.  The useful topologies include convergence
after smearing \(q_i\) and \(q_f\) with test functions, convergence after
analytic continuation to Euclidean time, and pointwise kernel convergence
under stronger regularity assumptions on \(V\).  Therefore the continuum
path-integral notation in this monograph always carries one of the following
meanings: a finite regulator, a strong operator limit, a smeared
distributional limit, a Euclidean measure or semigroup construction, or a
perturbative asymptotic expansion.

\section{Operator Ordering and Local Counterterms}

The function \(h(q,p)\) depends on the ordering convention used to represent
\(\widehat H\). For example, using left endpoints \(q_n\) in the short-time
matrix element and using right endpoints \(q_{n+1}\) define different symbols
for the same operator. Their difference is of order \(\hbar\) in the
continuum action and appears as a local counterterm in the time-sliced
description.

At fixed \(N\), two representative choices are
\[
  \cdots
  \bra{q_{n+1}}p_n\rangle
  \bra{p_n}\ee^{-\ii\epsilon\widehat H/\hbar}\ket{q_n}
  \cdots,
  \qquad
  \bra{p_n}\widehat H\ket{q_n}
  =
  h(q_n,p_n)\braket{p_n}{q_n},
\]
and
\[
  \cdots
  \bra{p_{n+1}}q_{n+1}\rangle
  \bra{q_{n+1}}\ee^{-\ii\epsilon\widehat H/\hbar}\ket{p_n}
  \cdots,
  \qquad
  \bra{q_n}\widehat H\ket{p_n}
  =
  h'(q_n,p_n)\braket{q_n}{p_n}.
\]
They discretize the kinetic term respectively as
\[
  \sum_{n=0}^{N-1}p_{n,a}(q_{n+1}^a-q_n^a),
  \qquad
  \sum_{n=0}^{N-1}p_{n+1,a}(q_{n+1}^a-q_n^a),
\]
after relabelling endpoints where necessary.  The two symbols \(h\) and
\(h'\) represent the same operator only with their specified ordering
conventions, and they can differ by terms of order \(\hbar\).  Such terms
survive in the exponent as local regulator-dependent additions to the
effective Lagrangian.

Thus the data of the regulated path integral are
\[
  (\text{partition},\ \text{measure},\ \text{operator symbol},\
  \text{boundary conditions}).
\]
A formula involving \(Dq\,Dp\) is meaningful as notation for a regulated
construction once these data have been declared.

\section{Lagrangian Form for Quadratic Momentum Dependence}

Assume the Hamiltonian symbol has the form
\[
  h(q,p)
  =
  \frac12\,G^{ab}(q)p_ap_b+V(q),
\]
where \(G^{ab}(q)\) is a positive-definite symmetric matrix and
\[
  G^{ab}(q)G_{bc}(q)=\delta^a_c.
\]
At fixed time slice, the momentum integral is Gaussian:
\[
  \int\frac{\dd^dp}{(2\pi\hbar)^d}
  \exp\left[
    \frac{\ii}{\hbar}
    \left(
      p_a\Delta q^a
      -\frac{\epsilon}{2}G^{ab}(q)p_ap_b
    \right)
  \right]
  =
  \frac{\sqrt{\det G_{ab}(q)}}{(2\pi\ii\hbar\epsilon)^{d/2}}
  \exp\left[
    \frac{\ii\epsilon}{\hbar}
    \frac12G_{ab}(q)v^av^b
  \right],
\]
where
\[
  \Delta q^a=q_{n+1}^a-q_n^a,
  \qquad
  v^a=\frac{\Delta q^a}{\epsilon}.
\]
The Lagrangian function obtained by Legendre transform is
\[
  L(q,\dot q)
  =
  \frac12G_{ab}(q)\dot q^a\dot q^b-V(q).
\]
Conversely, if a classical Lagrangian in generalized coordinates has this
metric form, as in the rigid-body configuration space example, then the
corresponding quantum Hamiltonian symbol is
\[
  h(q,p)=\frac12G^{ab}(q)p_ap_b+V(q)+O(\hbar)
\]
until an ordering convention and measure convention have been fixed.  The
\(O(\hbar)\) term records the same local regulator dependence described
above.
The finite-\(N\) configuration-space expression is therefore
\[
\begin{aligned}
  K_N(q_f,T;q_i,0)
  &=
  \int\prod_{n=1}^{N-1}\dd^dq_n
  \prod_{n=0}^{N-1}
  \frac{\sqrt{\det G_{ab}(q_n)}}{(2\pi\ii\hbar\epsilon)^{d/2}}
  \\
  &\quad\times
  \exp\left[
    \frac{\ii}{\hbar}
    \sum_{n=0}^{N-1}
    \epsilon\,
    L\left(q_n,\frac{q_{n+1}-q_n}{\epsilon}\right)
  \right].
\end{aligned}
\]
For \(G_{ab}=m\delta_{ab}\), the Lagrangian in this finite-dimensional
time-sliced kernel is
\[
  L(q,\dot q)=\frac{m}{2}\delta_{ab}\dot q^a\dot q^b-V(q).
\]

\section{Time-Ordered Insertions and Sources}

Let \(F(q(t_1),\ldots,q(t_r))\) be a product of coordinate insertions with
\[
  0<t_1<\cdots<t_r<T.
\]
At finite \(N\), choose the time slices nearest to the insertion times and
insert the corresponding coordinate functions into the finite-dimensional
integral. The resulting regulated expression represents
\[
  \bra{q_f}T\{\widehat q^{a_1}(t_1)\cdots
  \widehat q^{a_r}(t_r)\}U(T)\ket{q_i},
\]
where
\[
  \widehat q^a(t)
  =
  U(t)^{-1}\widehat q^aU(t)
\]
is the Heisenberg operator. The time ordering is inherited from the ordered
product of short-time evolution operators.

For a smooth compactly supported source \(J_a(t)\), define the finite-\(N\)
source factor by
\[
  \exp\left[
    \frac{\ii}{\hbar}
    \sum_{n=0}^{N-1}\epsilon\,J_a(t_n)q_n^a
  \right].
\]
The source-dependent kernel \(K_N[J]\) generates regulated time-ordered
coordinate insertions by differentiation with respect to \(J\). This is the
finite-dimensional ancestor of QFT generating functionals for Green
functions.

\section{Euclidean Traces and Periodic Boundary Conditions}

Assume that \(\widehat H\) is semibounded and that
\(\ee^{-\beta_{\mathrm T}\widehat H/\hbar}\) is trace class for
\(\beta_{\mathrm T}>0\). The Euclidean thermal trace is
\[
  Z(\beta_{\mathrm T})
  =
  \operatorname{Tr}_{\Hilb_{\mathcal Q}}
  \ee^{-\beta_{\mathrm T}\widehat H/\hbar}.
\]
Inserting the position resolution of the identity gives
\[
  Z(\beta_{\mathrm T})
  =
  \int_{\mathbb R^d}\dd^dq_0\,
  \bra{q_0}
  \ee^{-\beta_{\mathrm T}\widehat H/\hbar}
  \ket{q_0}.
\]
The same time-sliced construction as above, now with Euclidean time
\(\tau\in[0,\beta_{\mathrm T}]\), therefore has
\[
  q_N=q_0 .
\]
Thus the trace of a bosonic quantum-mechanical system is represented by the
Euclidean path integral over periodic paths
\[
  q(\tau+\beta_{\mathrm T})=q(\tau).
\]
This conclusion is a statement about the boundary condition produced by the
trace; the existence and convergence of the resulting continuum measure still
depend on the same operator and regularization hypotheses stated above.

\section{Vacuum Projection}

Assume \(\widehat H\) has a normalized ground state \(\ket{0}\), with energy
\(E_0\), and that a vector \(\ket\psi\in\Hilb_{\mathcal Q}\) satisfies
\(\braket{0}{\psi}\ne0\). If the spectrum above \(E_0\) is separated by a
gap, then
\[
  \lim_{\tau\to\infty}
  \frac{\ee^{-\tau(\widehat H-E_0)/\hbar}\ket\psi}
       {\left\|\ee^{-\tau(\widehat H-E_0)/\hbar}\ket\psi\right\|}
  =
  \ket{0}
\]
up to the phase of \(\braket{0}{\psi}\). Vacuum correlation functions can
therefore be obtained from long Euclidean-time evolution:
\[
  \bra{0}T\{\widehat q^{a_1}(t_1)\cdots
  \widehat q^{a_r}(t_r)\}\ket{0}
\]
is the normalized long-time limit of the corresponding transition amplitude
with insertions, after the Euclidean projection has been applied.

The next chapter will analyze the analytic continuation and Gaussian
integrals that make this construction computationally useful. The present
chapter has introduced the regulated path-integral data and its relation to
Hamiltonian time evolution.
