\chapter{Starting Data for Local Quantum Field Theory}
\label{ch:starting-data-local-qft}
\ConventionsMixed


\section{Spacetime And Operator Conventions}

Throughout the opening volume, \(D\) denotes the spacetime dimension. The
symbol \(\mathbb M^D\) denotes \(D\)-dimensional Minkowski affine space. After
a choice of inertial coordinates, \(\mathbb M^D\) is identified with
\(\mathbb R^D\). The metric tensor has components
\[
  \eta_{\mu\nu}=\operatorname{diag}(-1,1,\ldots,1),
\]
and units are chosen so that \(\hbar=c=1\). In the quantum-mechanical
path-integral derivations \(\hbar\) is displayed until the semiclassical
weighting and Wick rotation have been derived; after that point \(\hbar=1\)
is restored unless a dimensional check explicitly reintroduces it.  The
vacuum vector is denoted by \(\vac\); the same printed letter may occur with
subscripts for reconstructed cyclic vectors, but unsubscripted \(\vac\)
always denotes the vacuum of the theory under discussion. A vector
\(x\in\mathbb M^D\) is
spacelike when
\[
  x^2:=\eta_{\mu\nu}x^\mu x^\nu>0.
\]
A test function is a smooth compactly supported complex-valued function on
\(\mathbb M^D\), an element of \(C_c^\infty(\mathbb M^D)\), unless another
test-function space is explicitly declared. A pointlike quantum field
\(\widehat\Phi_A(x)\) is an operator-valued distribution: its primary
operator expression is the smeared field
\[
  \widehat\Phi_A(f)
  =
  \int_{\mathbb M^D}\dd^Dx\,f(x)\widehat\Phi_A(x),
  \qquad
  f\in C_c^\infty(\mathbb M^D).
\]
The integral sign is distributional notation for evaluating the field on the
test function \(f\).  When \(\widehat\Phi_A(f)\) is unbounded, its common dense
domain is part of the field framework being used.  The symbol
\(\mathcal B(\Hilb)\) denotes the algebra of bounded linear operators on a
Hilbert space \(\Hilb\), and \(\mathcal U(\Hilb)\) denotes its unitary group.
The label \(A\) denotes a component in a finite-dimensional representation
space for Lorentz and internal indices, or a composite-operator label, as
specified in the chapter where the field is introduced.

This notation distinguishes mathematical object classes.  In Hilbert-space
formulas a hat on a field symbol denotes an operator-valued distribution:
\(\widehat\phi\) for a scalar field and \(\widehat\Phi_A\) for a multiplet.
Unhatted fields such as \(\phi\), \(\varphi\), or \(\Phi_E\) denote classical
variables, path-integral variables, or Euclidean Schwinger/random fields unless
a chapter explicitly declares a different use.  Fourier transforms of such
variables are written with a tilde, for example \(\widetilde\phi(k)\).

The test-function space is part of the framework data.  Locality statements
use \(C_c^\infty\) because support controls localization in bounded spacetime
regions.  Spectral, Wightman, and Euclidean reconstruction statements often
use the Schwartz space \(\mathcal S\), because temperedness and Fourier
analysis are then built into the topology.  A chapter may use either declared
nuclear test-function space, provided all field and correlation distributions
are continuous for that topology.  A tempered field on \(\mathcal S\) restricts
to compactly supported test functions for local statements; a compactly
supported-test distribution is treated as tempered only after a stated growth
or continuity hypothesis supplies a continuous extension to \(\mathcal S\).

Euclidean correlation distributions are called Schwinger functions when they
are being used as reconstruction data and are then denoted \(S_n\).  In
path-integral perturbation chapters the notation \(G_{E,n}\) denotes the same
kind of Euclidean \(n\)-point distribution before the OS reconstruction
hypotheses have been imposed; \(G_E\) without an \(n\)-index is reserved for a
two-point kernel or its Fourier transform.

The connected proper orthochronous Lorentz group is denoted
\(SO^+(1,D-1)\), and the connected Poincare group is
\[
  \Poinc=\mathbb R^D\rtimes SO^+(1,D-1).
\]
Two subsets
\(\mathcal O_1,\mathcal O_2\subset\mathbb M^D\) are spacelike separated when
\((x-y)^2>0\) for every \(x\in\mathcal O_1\) and
\(y\in\mathcal O_2\).

\section{Local Continuum Theories And Effective Presentations}

The organizing structure of quantum field theory is quantum dynamics together
with spacetime localization.  The word ``localization'' has a precise content:
observables, or field coordinates for observables, are attached to spacetime
regions or test functions; spacetime symmetries move these localized objects
covariantly; and observables localized in spacelike separated regions obey the
appropriate commutativity law.  The exact algebraic, field-theoretic, or
Euclidean form of this structure must always be specified.

Two different mathematical statuses will appear throughout the monograph.
They share locality, but they are not the same assertion.

\begin{definition}[Continuum local quantum field theory]
  A continuum local quantum field theory is a local quantum theory whose
  declared correlation functions, local algebras, or Hilbert-space operators
  exist after the ultraviolet regulator has been removed, in a stated
  topology and within a stated spacetime and state-sector framework.  In a
  Minkowski vacuum presentation this includes Poincare symmetry, a spectrum
  condition, local observables or local field distributions, and
  microcausality.  The phrase ``infinitely many degrees of freedom'' will be
  used in the operational sense that the local data are indexed by
  infinite-dimensional test-function or region data and are not specified by a
  finite list of canonical coordinates.
\end{definition}

\begin{definition}[Effective field theory presentation]
  An effective field theory presentation is a regulated or renormalized local
  prescription for computing observables in a specified range of length,
  energy, and external-state scales.  Its defining data include the fields or
  local variables, the symmetries, the regulator or renormalization
  prescription, the expansion parameters, and the class of observables for
  which the expansion is claimed.  When it is perturbative, it defines an
  asymptotic expansion of Green functions, and of scattering quantities only
  after the nonperturbative scattering framework and LSZ interpretation have
  been supplied.
\end{definition}

Power-counting renormalizability is a property of an effective presentation
near a chosen fixed point or scaling regime: only finitely many local
operators are needed as independent counterterms at a fixed order in the
declared expansion.  It is not, by itself, a proof that the ultraviolet
continuum theory exists.  Power-counting nonrenormalizable presentations are
also local effective theories when the infinitely many allowed operators are
organized by scale and symmetry.

\begin{remark}[Conformal perturbation as a controlled construction method]
\label{rem:opening-cft-perturbation-scope}
A deformation of a conformal field theory is a precise construction problem
only after the conformal theory has already been supplied as local QFT data.
The input consists of a stress tensor, a Hilbert space or Euclidean
reconstruction datum, local operators \(O_A\) with specified normalization and
contact-term convention, and a regulator or subtraction prescription for
insertions of
\[
  \exp\!\left(-\sum_A g^A\int\dd^Dx\,O_A(x)\right).
\]
When these data exist and the deformation parameters are controlled, conformal
perturbation theory produces a local expansion of deformed correlation
functions and beta functions.  Its hypotheses are therefore:
\begin{enumerate}
  \item an already constructed or otherwise supplied fixed point;
  \item identified operator insertions, including their coincident-point
  renormalization;
  \item a domain in coupling space where the expansion or nonperturbative
  deformation is defined;
  \item a reconstruction or regulator argument connecting the deformed
  correlators to local QFT data.
\end{enumerate}
Without these inputs, the phrase ``perturb a UV CFT'' is only a description
of a possible route.  Asymptotically free gauge theories, constructive scalar
models, lattice scaling limits, stochastic-quantization constructions, and
effective presentations of the Standard Model may each possess local QFT
content in their own frameworks; none is thereby defined by conformal
perturbation theory.  This monograph uses CFT perturbations when their
operator-algebra and regulator hypotheses are declared, for example in
Wilson--Fisher and two-dimensional integrable-flow problems, and does not use
them as the definition of quantum field theory.
\end{remark}

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      root/.style={draw,rounded corners=4pt,line width=0.85pt,
        fill=black!3,minimum width=5.1cm,minimum height=0.7cm,align=center},
      box/.style={draw,rounded corners=4pt,line width=0.8pt,
        minimum width=5.4cm,minimum height=1.0cm,align=center},
      textbox/.style={draw,rounded corners=4pt,line width=0.55pt,
        align=left,inner sep=5pt,text width=5.8cm},
      arrow/.style={-{Latex[length=2mm]},line width=0.8pt}
    }
    \node[root] (root) at (0,2.35)
      {quantum dynamics with spacetime localization};
    \node[box,fill=blue!5] (local) at (-4.0,0.82)
      {continuum local QFT\\data exist after removing the ultraviolet regulator};
    \node[box,fill=orange!7] (eft) at (4.0,0.82)
      {effective field theory presentation\\local expansion valid in a declared scale window};
    \draw[arrow] (root.south west) -- (local.north);
    \draw[arrow] (root.south east) -- (eft.north);

    \node[textbox,anchor=north,fill=blue!2] at (-4.0,-0.15)
      {\(\Hilb\), Poincare representation, spectrum condition\\
       local observables or local field distributions\\
       microcausality for spacelike separated supports\\[2pt]
       guiding examples: low-dimensional constructive scalar theories;
       scaling limits of lattice models near criticality; Yang--Mills/QCD as
       physical continuum local gauge theories};

    \node[textbox,anchor=north,fill=orange!3] at (4.0,-0.15)
      {regulated path integral or equivalent local prescription\\
       symmetry-controlled local operator expansion\\
       long-distance or low-energy observables in the stated regime\\[2pt]
       examples: \(\phi^4\) theory in \(D=4\), QED, and the Standard Model as
       power-counting renormalizable perturbative presentations; chiral
       perturbation theory and perturbative quantum gravity as
       nonrenormalizable effective presentations};
  \end{tikzpicture}
  }
  \caption{The opening distinction between continuum local QFT data and
  effective field theory presentations.  Both sides use locality; the
  difference is the mathematical status of the construction and the range of
  claims made from it.}
  \label{fig:opening-local-qft-eft}
\end{figure}

\section{Opening Framework}

The first construction takes place in a flat-spacetime vacuum sector. The
following definition gives the data used in the opening volume.

\begin{definition}[Vacuum Minkowski local quantum theory]
\label{def:vacuum-minkowski-local-qft}
In the opening framework, a vacuum Minkowski local quantum theory consists of:
\begin{enumerate}
  \item a complex Hilbert space \(\Hilb\);
  \item a spacetime symmetry convention.  Let \(G_{\rm Lor}\) be
  \(SO^+(1,D-1)\), or the appropriate double cover when spinor fields are
  included, and let
  \(\pi:G_{\rm Lor}\to SO^+(1,D-1)\) be the covering projection.  The connected
  Poincare group used here is
  \[
    \mathcal P=\mathbb R^D\rtimes G_{\rm Lor},
    \qquad
    (a,\Lambda)(b,\Gamma)
    =
    (a+\pi(\Lambda)b,\Lambda\Gamma).
  \]
  It acts on \(\mathbb M^D\) from the left by
  \[
    g x=\pi(\Lambda)x+a,
    \qquad g=(a,\Lambda)\in\mathcal P,
  \]
  and \(g\mathcal O:=\{gx:x\in\mathcal O\}\) for a region
  \(\mathcal O\subset\mathbb M^D\);
  \item a strongly continuous unitary representation
  \[
    U:\mathcal P\to\mathcal U(\Hilb),
    \qquad
    g\mapsto U(g),
  \]
  as a left representation, \(U(g_1g_2)=U(g_1)U(g_2)\), where
  \(\mathcal U(\Hilb)\) is the unitary group of \(\Hilb\);
  \item self-adjoint translation generators \(P^\mu\), equivalently
  \(P_\mu=\eta_{\mu\nu}P^\nu\), defined by
  \[
    U(a,1)=\exp(\ii a_\mu P^\mu)=\exp(\ii a^\mu P_\mu).
  \]
  Thus a positive time translation gives
  \(\exp(-\ii a^0P^0)\), with \(P^0\) the Hamiltonian.  Let \(E_P\) be the
  joint projection-valued spectral measure of the contravariant
  energy-momentum \(P=(P^0,\ldots,P^{D-1})\).  The spectrum condition is
  \[
    \operatorname{supp}E_P\subset
    \overline V_+
    =
    \{p\in\mathbb R^D:p^0\ge0,\ \eta_{\mu\nu}p^\mu p^\nu\le0\};
  \]
  \item a Poincare-invariant unit vector \(\vac\in\Hilb\),
  \[
    U(g)\vac=\vac,\qquad g\in\mathcal P.
  \]
  In this pure-vacuum-sector framework, vacuum uniqueness is part of the
  datum: the zero-momentum spectral projection is one-dimensional,
  \[
    E_P(\{0\})\Hilb
    =
    \mathbb C\vac;
  \]
  \item an assignment
  \[
    \mathcal O\longmapsto\Obs(\mathcal O)\subset\mathcal B(\Hilb)
  \]
  from bounded open regions \(\mathcal O\subset\mathbb M^D\) to unital
  concrete \(*\)-subalgebras of the bounded operators on \(\Hilb\), satisfying
  covariance, isotony, and locality:
  \[
    U(g)^{-1}\Obs(\mathcal O)U(g)
    =
    \Obs(g\mathcal O),
    \qquad g\in\mathcal P,
  \]
  \[
    \mathcal O_1\subset\mathcal O_2
    \quad\Longrightarrow\quad
    \Obs(\mathcal O_1)\subset\Obs(\mathcal O_2),
  \]
  and
  \[
    [A,B]=0
    \quad
    \text{for }A\in\Obs(\mathcal O_1),\ B\in\Obs(\mathcal O_2)
  \]
  whenever \(\mathcal O_1\) and \(\mathcal O_2\) are spacelike separated, with
  the graded commutator
  \[
    [A,B]_{\rm gr}=AB-(-1)^{|A||B|}BA
  \]
  used for homogeneous odd operators after a fermion parity grading and the
  parities \(|A|,|B|\in\mathbb Z/2\mathbb Z\) have been specified;
  \item vacuum cyclicity for the bounded local observable content.  Let
  \(\Obs_{\rm loc}^{\rm alg}\) denote the unital \(*\)-algebra generated in
  \(\mathcal B(\Hilb)\) by \(\bigcup_{\mathcal O}\Obs(\mathcal O)\), where the
  union is over bounded open regions.  Then
  \[
    \overline{\Obs_{\rm loc}^{\rm alg}\vac}^{\,\|\cdot\|_{\Hilb}}
    =
    \Hilb;
  \]
  \item when unbounded field coordinates are part of the presentation, a
  specified field-domain datum
  \[
    (\mathcal D_\Phi,\{\widehat\Phi_A\}_{A\in I}).
  \]
  Here \(I\) is the field-label set, \(\mathcal D_\Phi\subset\Hilb\) is a dense
  linear subspace with \(\vac\in\mathcal D_\Phi\), \(U(g)\mathcal D_\Phi\subset
  \mathcal D_\Phi\) for all \(g\in\mathcal P\), and each smeared field is a
  linear operator
  \[
    \widehat\Phi_A(f):\mathcal D_\Phi\to\mathcal D_\Phi,
    \qquad f\in\mathcal T_A,
  \]
  where \(\mathcal T_A\) is the declared test-function space for that field
  component.  Finite products of such smeared fields are defined on
  \(\mathcal D_\Phi\).  If adjoints are used, then
  \(\mathcal D_\Phi\subset\operatorname{Dom}(\widehat\Phi_A(f)^*)\) and the adjoint
  fields preserve \(\mathcal D_\Phi\).  For every
  \(\psi,\chi\in\mathcal D_\Phi\), the matrix elements of finite smeared
  products are continuous multilinear functionals of their test functions.
  Field covariance, field locality, and any relation between
  \(\widehat\Phi_A(f)\) with \(\operatorname{supp}f\subset\mathcal O\) and
  \(\Obs(\mathcal O)\) are imposed on this same domain or by a specified
  affiliation to the chosen local observable algebra.
\end{enumerate}
\end{definition}

The uniqueness clause is a spectral assumption on the chosen representation.
It identifies the projection onto translation-invariant vectors with the
rank-one projection
\(|\vac\rangle\langle\vac|\).  Consequently, after subtracting the vacuum
expectation value of a scalar local field, the vector
\(\widehat\phi(f)\vac\) has no residual \(p=0\) component.  This is the
assumption used in the K\"all\'en--Lehmann analysis when the
\((\vac,\cdot)\) projection removes the vacuum atom before the invariant-mass
decomposition is formed.  If a theory has several infinite-volume vacua or a
mixed representation of phases, one must first choose a pure vacuum sector or
else keep the additional \(p=0\) spectral summands explicitly; they sit at
\(\mu^2=0\) as zero-momentum atoms, distinct from massless one-particle states
whose momenta lie on the nonzero light cone.

The covariance convention has the following infinitesimal form.  With
\(U(a,1)=\exp(\ii a^\mu P_\mu)=\exp(\ii a_\mu P^\mu)\), a
coordinate-labelled scalar field obeys
\[
  U(a,1)^{-1}\widehat\phi(x)U(a,1)
  =
  \widehat\phi(x+a),
  \qquad
  [P_\mu,\widehat\phi(x)]
  =
  \ii\,\partial_\mu\widehat\phi(x)
\]
as an identity of distributions on a common invariant domain.  Equivalently,
conjugation by \(U(a,1)\) shifts the coordinate label by \(-a\).  The local-net
covariance law above uses the same passive coordinate convention: conjugating
an observable by \(U(g)^{-1}\), with \(g=(a,\Lambda)\), sends its localization
region to \(g\mathcal O=\pi(\Lambda)\mathcal O+a\).

In this opening convention \(\Obs(\mathcal O)\) is a concrete algebra of
bounded observables on the chosen Hilbert space and is not assumed to be norm
closed or weakly closed unless this is stated.  If a \(C^*\)-algebra or von
Neumann algebra is needed, the relevant closure or bicommutant is included in
the datum or constructed explicitly.  This convention is sufficient for the
first dependency layer: localization, covariance, the spectrum condition,
vacuum matrix elements, and comparison with field coordinates.

This framework describes a vacuum representation on flat spacetime. Thermal
states, boundary conditions, curved spacetimes, and non-vacuum superselection
sectors are introduced by changing the stated data: the state space, spacetime
category, covariance group, region assignment, or representation sector.
Existing Wightman, Euclidean, algebraic, constructive, perturbative, and
factorization frameworks will be introduced as precise frameworks with their
own domains, beginning with the Wightman and algebraic formulations.

\begin{openproblem}[Four-dimensional constructive local QFT]
\label{op:four-dimensional-constructive-qft}
A constructive solution of a four-dimensional local quantum field theory is a
construction of the correlation functions, Hilbert space, local fields or
observable net, symmetry representation, and positivity structure in a topology
strong enough to verify the stated framework.  In a Wightman presentation this
means tempered vacuum distributions satisfying covariance, the spectrum
condition, locality, positivity, and a reconstruction theorem.  In an
Osterwalder--Schrader presentation it means Euclidean Schwinger distributions
with reflection positivity, Euclidean covariance, symmetry, clustering or the
declared vacuum-sector condition, and the regularity hypotheses needed for
OS reconstruction.  In either presentation, an interacting model means that the
truncated correlation functions and local operator content are not those of a
Gaussian or generalized-free field alone, and that the local stress-tensor,
charge, or gauge-invariant observables required in the later physical
discussion are part of the constructed data.

Completed constructive examples occupy special regimes: \(P(\phi)_2\) models,
including \(\phi^4_2\), and massive \(\phi^4_3\) give non-Gaussian scalar
models after their model-specific Wick ordering and local renormalizations;
other superrenormalizable low-dimensional scalar--fermion models are
constructed under similarly explicit hypotheses; and two-dimensional
Yang--Mills has rigorous gauge-theory constructions through holonomy and
heat-kernel measures.  Four-dimensional physics presents a different
constructive problem.  Free and generalized-free Wightman fields are available,
but the usual interacting scalar and gauge theories require separate
continuum-limit theorems.  Standard positive \(\lambda\phi^4_D\) scaling-limit
families in \(D\geq4\) are constrained by triviality theorems, with the
four-dimensional marginal case governed by the Aizenman--Duminil-Copin
triviality theorem for the standard Ising/\(\phi^4_4\) scaling problem.
Four-dimensional pure Yang--Mills, QCD, and other asymptotically free gauge
theories still require a construction of the continuum gauge-invariant
Schwinger functions or Wightman theory with the desired spectral properties.
Perturbative renormalizability and asymptotic freedom are mechanisms inside a
specified expansion; they are not by themselves constructive existence
theorems for the nonperturbative continuum theory.
\end{openproblem}

\section{Framework Presentations And Comparison Maps}

The opening data admit several mathematically precise presentations.  Each
presentation emphasizes a different class of objects, and each comparison map
carries hypotheses.

\begin{definition}[Wightman field presentation]
  A Wightman field presentation consists of Hilbert-space data
  \[
    (\Hilb,\vac,U,\mathcal D,\{\widehat\Phi_A\}_{A\in I})
  \]
  where \(I\) is an index set for fields, \(\mathcal D\subset\Hilb\) is a
  dense common invariant domain containing \(\vac\), and each \(\widehat\Phi_A\) is an
  operator-valued distribution on a declared invariant test-function space
  \(\mathcal T\), with common domain \(\mathcal D\).  In the standard
  Wightman case \(\mathcal T=\mathcal S(\mathbb M^D)\), so the vacuum
  distributions below are tempered.  The domain assumptions are
  \[
    \vac\in\mathcal D,\qquad
    U(g)\mathcal D\subset\mathcal D\quad(g\in\mathcal P),\qquad
    \widehat\Phi_A(f)\mathcal D\subset\mathcal D\quad(A\in I,\ f\in\mathcal T),
  \]
  and, for every \(n\ge1\),
  \[
    \widehat\Phi_{A_1}(f_1)\cdots\widehat\Phi_{A_n}(f_n)\mathcal D
    \subset
    \mathcal D,
    \qquad A_i\in I,\quad f_i\in\mathcal T.
  \]
  Matrix elements of finite products are required to be continuous multilinear
  functionals of the test functions:
  \[
    (f_1,\ldots,f_n)
    \longmapsto
    \left\langle\psi,
    \widehat\Phi_{A_1}(f_1)\cdots\widehat\Phi_{A_n}(f_n)\chi
    \right\rangle
    \in\mathbb C,
    \qquad \psi,\chi\in\mathcal D .
  \]

  Let \(g=(a,\Lambda)\in\mathcal P\), \(gx=\pi(\Lambda)x+a\), and
  \(f_g(y)=f(g^{-1}y)\).  If the fields carry a finite-dimensional Lorentz
  representation \(R\) of \(G_{\rm Lor}\), covariance is the smeared identity
  \[
    U(g)^{-1}\widehat\Phi_A(f)U(g)
    =
    R(\Lambda)_A{}^B\,\widehat\Phi_B(f_g),
  \]
  with summation over repeated component labels.  In point-field notation this
  is
  \[
    U(g)^{-1}\widehat\Phi_A(x)U(g)=R(\Lambda)_A{}^B\widehat\Phi_B(gx).
  \]
  The representation \(U\) satisfies the spectrum condition of
  Definition~\ref{def:vacuum-minkowski-local-qft}: if \(E_P\) is the joint
  spectral measure of the translation generators, then
  \[
    \operatorname{supp}E_P\subset\overline V_+.
  \]

  Locality is imposed after smearing:
  \[
    [\widehat\Phi_A(f),\widehat\Phi_B(h)]_{\rm gr}=0
  \]
  whenever \(\operatorname{supp}f\) and \(\operatorname{supp}h\) are spacelike
  separated, with the ordinary commutator in the purely bosonic case.

  Its primary correlation data are the vacuum distributions whose distribution
  kernels are written
  \[
    W_{A_1\cdots A_n}(x_1,\ldots,x_n)
    =
    \bra{\vac}
    \widehat\Phi_{A_1}(x_1)\cdots\widehat\Phi_{A_n}(x_n)
    \ket{\vac}.
  \]
  Equivalently,
  \[
    W_{A_1\cdots A_n}(f_1,\ldots,f_n)
    =
    \left\langle\vac,
    \widehat\Phi_{A_1}(f_1)\cdots\widehat\Phi_{A_n}(f_n)\vac
    \right\rangle .
  \]
  Translation invariance gives a distribution in difference variables.  With
  \[
    \xi_j=x_{j+1}-x_j,\qquad
    \mathcal W_{A_1\cdots A_n}(\xi_1,\ldots,\xi_{n-1})
    =
    W_{A_1\cdots A_n}
    (0,\xi_1,\xi_1+\xi_2,\ldots,\xi_1+\cdots+\xi_{n-1}),
  \]
  the Wightman spectral condition is the support equation
  \[
    \operatorname{supp}
    \widehat{\mathcal W}_{A_1\cdots A_n}
    \subset
    \overline V_+^{\,n-1}.
  \]
  Positivity of the hierarchy means the following.  For every finite sequence
  of test tensors in the declared test-function space
  \[
    F=\{F^{(n)}_{A_1\cdots A_n}\}_{0\leq n\leq N},
    \qquad
    F^{(0)}\in\mathbb C,
  \]
  define
  \[
    \Psi_F
    =
    F^{(0)}\vac
    +
    \sum_{n=1}^N
    \sum_{A_1,\ldots,A_n}
    \int F^{(n)}_{A_1\cdots A_n}(x_1,\ldots,x_n)
    \widehat\Phi_{A_1}(x_1)\cdots\widehat\Phi_{A_n}(x_n)\vac\,
    \dd^Dx_1\cdots\dd^Dx_n .
  \]
  The condition is \(\langle\Psi_F,\Psi_F\rangle\geq0\), with the inner product
  evaluated by the distributions \(W_n\):
  \[
    \langle\Psi_F,\Psi_F\rangle\ge0.
  \]
  Vacuum cyclicity is the density equation
  \[
    \overline{
    \operatorname{span}\{
    \widehat\Phi_{A_1}(f_1)\cdots\widehat\Phi_{A_n}(f_n)\vac
    : n\ge0,\ f_i\in\mathcal T,\ A_i\in I
    \}
    }^{\,\|\cdot\|_{\Hilb}}
    =
    \Hilb .
  \]
\end{definition}

\begin{definition}[Local-net presentation]
  A local-net presentation consists of a specified class
  \(\mathcal K(\mathbb M^D)\) of open spacetime regions, a quasilocal
  \(*\)-algebra or \(C^*\)-algebra \(\Obs_{\rm qloc}\), and an assignment
  \[
    \mathcal O\longmapsto\Obs(\mathcal O)\subset\Obs_{\rm qloc}
  \]
  for \(\mathcal O\in\mathcal K(\mathbb M^D)\).  The assignment satisfies
  isotony, locality, covariance under the declared spacetime symmetry group,
  and any additional structural conditions imposed in the problem under
  study, such as additivity, Haag duality, or the time-slice property.  A
  state is a normalized positive linear functional
  \(\omega:\Obs_{\rm qloc}\to\mathbb C\); it gives a Hilbert-space
  representation by the GNS construction.
\end{definition}

\begin{definition}[Euclidean correlation presentation]
  Let \(E=\mathbb R^D\) be Euclidean space and let \(V\) be the
  finite-dimensional complex field-label space, with homogeneous basis labels
  \(a\) and parity \(|a|\in\mathbb Z/2\mathbb Z\) when a grading is present.  A
  Euclidean correlation presentation consists of a test-function space
  \(\mathcal T_E\), usually \(\mathcal S(E)\) in the tempered OS framework, an
  adjoint antilinear involution \(a\mapsto a^\dagger\) on labels, a Euclidean
  rotation representation \(R_E\) on \(V\), and a hierarchy
  \[
    S_n\in\mathcal T_{E,\circ}'(E^n;(V^*)^{\otimes n}),
    \qquad n\ge0,\qquad S_0=1.
  \]
  Here \(\mathcal T_{E,\circ}(E^n;V^{\otimes n})\) denotes the corresponding
  \(V^{\otimes n}\)-valued zero-diagonal test-function space on \(E^n\): test
  functions vanish with all derivatives on every partial diagonal
  \(x_i=x_j\).  The prime denotes its continuous dual, and \(S_n(F)\) denotes
  the distributional pairing.  A contact-term extension of \(S_n\) from
  \(\mathcal T_{E,\circ}\) to the full \(\mathcal T_E\) is additional
  renormalization data and is not required for OS reconstruction.
  The component notation
  \[
    S_{a_1\cdots a_n}(x_1,\ldots,x_n)
  \]
  is kernel notation for the distributional pairing of \(S_n\) with
  \(V^{\otimes n}\)-valued test functions.  If
  \(g=(b,r)\in\mathbb R^D\rtimes O(D)\), \(gx=rx+b\), define the action on test
  tensors by
  \[
    (gF)^{a_1\cdots a_n}(x_1,\ldots,x_n)
    =
    R_E(r)^{a_1}{}_{b_1}\cdots R_E(r)^{a_n}{}_{b_n}\,
    F^{b_1\cdots b_n}(g^{-1}x_1,\ldots,g^{-1}x_n).
  \]
  Euclidean covariance is the distributional identity
  \[
    S_n(gF)=S_n(F),
    \qquad g\in\mathbb R^D\rtimes O(D).
  \]
  For a permutation \(\sigma\in\mathfrak S_n\), graded permutation symmetry is
  \[
    S_{a_{\sigma(1)}\cdots a_{\sigma(n)}}
    (x_{\sigma(1)},\ldots,x_{\sigma(n)})
    =
    \epsilon(\sigma;a_1,\ldots,a_n)\,
    S_{a_1\cdots a_n}(x_1,\ldots,x_n),
  \]
  as an equality of distributions, where
  \[
    \epsilon(\sigma;a_1,\ldots,a_n)
    =
    \prod_{\substack{i<j\\ \sigma(i)>\sigma(j)}}
    (-1)^{|a_i||a_j|}.
  \]
  In the ungraded case all parities are zero.  Hermiticity is
  \[
    \overline{
    S_{a_1\cdots a_n}(x_1,\ldots,x_n)}
    =
    S_{a_n^\dagger\cdots a_1^\dagger}(x_n,\ldots,x_1),
  \]
  again as a distributional statement.

  Choose Euclidean coordinates
  \(x=(\tau,\mathbf x)\) and let
  \[
    \theta(\tau,\mathbf x)=(-\tau,\mathbf x)
  \]
  be time reflection.  Set \(E_+=\{(\tau,\mathbf x)\in E:\tau>0\}\).  Reflection
  positivity is the requirement that for every finite sequence of test tensors
  \[
    F=\{F^{(n)}_{a_1\cdots a_n}\}_{0\leq n\leq N},
    \qquad
    F^{(n)}\in\mathcal T_E(E^n;V^{\otimes n}),
    \qquad
    \operatorname{supp}F^{(n)}
    \subset E_+^n,
  \]
  the quadratic form
  \[
  \begin{aligned}
    \langle F,F\rangle_{\rm OS}
    &=
    \sum_{m,n=0}^N
    \sum_{\substack{a_1,\ldots,a_m\\ b_1,\ldots,b_n}}
    \int
    \overline{F^{(m)}_{a_1\cdots a_m}(x_1,\ldots,x_m)}
    F^{(n)}_{b_1\cdots b_n}(y_1,\ldots,y_n)
    \\
    &\qquad\qquad\qquad\quad
    \times
    S_{a_m^\dagger\cdots a_1^\dagger b_1\cdots b_n}
    (\theta x_m,\ldots,\theta x_1,y_1,\ldots,y_n)
    \prod_{i=1}^m\dd^Dx_i\prod_{j=1}^n\dd^Dy_j
  \end{aligned}
  \]
  is nonnegative.  The convention \(S_0=1\) is used in the terms with
  \(m=0\) or \(n=0\).

  The regularity and growth condition is also part of the data.  The corrected
  OS-II reconstruction theorem requires more than ordinary temperedness at each
  fixed \(n\): one needs uniform control on the growth of the distributional
  order with \(n\).  In a simple full-Schwartz presentation this includes
  constants \(A,B,\gamma>0\) and a seminorm order \(N(n)\) growing at most
  linearly in \(n\), such that every
  \(F\in\mathcal S_\circ(E^n;V^{\otimes n})\) obeys a Schwartz-seminorm bound,
  where \(\mathcal S_\circ\) denotes the Schwartz subspace of functions flat on
  the partial diagonals,
  using any fixed norm on the finite-dimensional space \(V^{\otimes n}\):
  \[
    |S_n(F)|
    \le
    A\,B^n(n!)^\gamma
    \sum_{|\alpha|,|\beta|\le N(n)}
    \sup_{X\in E^n}
    \left|X^\beta\partial^\alpha F(X)\right|.
  \]
  If a different nuclear test-function space is chosen, the corresponding
  continuity seminorms and the growth assumptions needed for OS reconstruction
  must be stated with the presentation.

  Under the full Osterwalder--Schrader hypotheses,
  quotienting by the null space of this form and completing reconstructs the
  Lorentzian Hilbert space; analytic continuation then gives Wightman boundary
  values.
\end{definition}

The comparison from fields to a local net is obtained, when the needed
operator-domain and affiliation statements hold, by assigning to a region
\(\mathcal O\) the von Neumann algebra generated by bounded functions,
resolvents, or other bounded functional-calculus representatives of smeared
fields \(\widehat\Phi_A(f)\) with \(\operatorname{supp}f\subset\mathcal O\).  The
comparison from a local net to a Hilbert-space theory is the GNS construction
applied to a state, followed by the representation of the local algebras on
the GNS Hilbert space.  Recovering point-field coordinates from a represented
net requires additional phase-space, regularity, or affiliated-field
hypotheses; it is not a formal consequence of the net axioms alone.  The
comparison from Euclidean correlations to Lorentzian fields is the corrected
OS-II reconstruction theorem followed by analytic continuation to Wightman
boundary values.  Conversely, Wightman fields give Euclidean Schwinger
functions by tube analyticity and restriction to noncoincident imaginary-time
configurations.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      box/.style={draw,rounded corners=4pt,line width=0.8pt,fill=blue!4,
        text width=3.15cm,minimum height=0.95cm,inner sep=4pt,align=center},
      pbox/.style={draw,rounded corners=4pt,line width=0.8pt,fill=orange!7,
        text width=3.15cm,minimum height=0.95cm,inner sep=4pt,align=center},
      arrow/.style={-{Latex[length=2mm]},line width=0.85pt},
      darrow/.style={-{Latex[length=2mm]},dashed,line width=0.85pt},
      note/.style={align=center,font=\scriptsize}
    }
    \node[box] (w) at (-4.45,1.70)
      {Wightman fields\\\(\widehat\Phi_A(f),\,W_n\)};
    \node[box] (a) at (4.45,1.70)
      {local nets\\\(\mathcal O\mapsto\Obs(\mathcal O)\)};
    \node[box] (e) at (0,-2.10)
      {OS Euclidean data\\\(S_n\)};
    \node[pbox] (p) at (0,-4.15)
      {regulated path integral\\or lattice system};
    \node[box,fill=green!7] (h) at (0,0.20)
      {Hilbert-space vacuum sector\\\((\Hilb,\vac,U)\)};
    \draw[arrow] (w) to[bend left=10]
      node[midway,above] {closures and affiliation} (a);
    \draw[arrow] (a) to[bend left=10]
      node[midway,below] {field coordinates} (w);
    \draw[arrow] (e) to[bend left=12]
      node[pos=0.46,left] {OS-II reconstruction} (w);
    \draw[arrow] (w) to[bend left=12]
      node[pos=0.48,right] {tube restriction} (e);
    \draw[arrow] (a.south west) to[out=-150,in=25]
      node[pos=0.55,above,sloped] {GNS} (h.east);
    \draw[arrow] (h.west) to[out=155,in=-25]
      node[pos=0.50,below,sloped] {cyclic field domains} (w.south east);
    \draw[darrow] (p)--(e)
      node[midway,right] {continuum Schwinger limits};
    \node[note] at (0,-5.20)
      {Each arrow includes hypotheses; the comparison data are part of the theory being used.};
  \end{tikzpicture}
  }
  \caption{Three nonperturbative presentations of local QFT and the principal
  comparison maps among them.}
  \label{fig:opening-framework-comparison}
\end{figure}

The theorem-level comparison register is as follows.
\begin{itemize}[leftmargin=*]
\item \textbf{Wightman hierarchy to Wightman fields.}
Theorem~\ref{thm:wightman-reconstruction} reconstructs the cyclic field
presentation from a positive Wightman hierarchy with covariance, spectrum,
Hermiticity, locality, and tempered continuity.
\item \textbf{Wightman fields to AQFT.}
Theorem~\ref{thm:wightman-to-aqft-net} constructs a represented local net once
the relevant smeared Hermitian fields have essentially self-adjoint closures
and strong locality of their spectral projections.
\item \textbf{AQFT to Wightman fields.}
Theorem~\ref{thm:aqft-to-wightman-fields} records the converse comparison:
a represented positive-energy local net gives Wightman field coordinates only
after affiliated operator-valued tempered distributions on a common invariant
domain have been supplied.
\item \textbf{OS data to Wightman fields.}
Theorem~\ref{thm:os-reconstruction} reconstructs a Lorentzian presentation
from OS-admissible Euclidean data.  Its analytic input is the OS-II
positive-time semigroup and linear-growth package, not reflection positivity
alone.
\item \textbf{Path integral to Wightman fields.}
Corollary~\ref{cor:path-integral-schwinger-limit-to-wightman} gives the
composition: a regulated path integral determines a Wightman presentation only
after its continuum Schwinger limits exist and satisfy the OS hypotheses.
\item \textbf{Wightman fields to OS data.}
Tube analyticity from the Wightman spectrum condition gives Euclidean
Schwinger functions by restriction to noncoincident imaginary-time
configurations; this is the reverse Wick-rotation direction discussed in
Chapter~\ref{ch:volume-iv-wightman-fields-reconstruction}.
\end{itemize}

These definitions make the foundational objects explicit.  A later
calculation may use fields, path integrals, local algebras, or Euclidean
correlators; the framework in which the calculation lives and the comparison
maps needed to interpret it must be stated.

\section{Status of Axiom Systems and Constructive Examples}
\label{sec:status-axiom-systems-constructive-examples}

The presentations above have a precise logical role.  They are theorem-bearing
frameworks for local quantum field theory: each one specifies objects,
topologies, positivity conditions, covariance conditions, and reconstruction
or comparison theorems.  A result proved in one framework has the hypotheses
of that framework as part of its statement.  Thus a theorem proved for
Wightman fields is a theorem for theories admitting the declared Wightman
presentation; a theorem proved for a represented local net is a theorem under
the local-net hypotheses and any additional phase-space or modular hypotheses
used in the proof; and a theorem proved from OS data is a theorem after the
OS-II regularity and analytic-continuation input has been verified.

The physical expectation is broader and more constructive.  A
Poincare-invariant ultraviolet-complete local QFT in a vacuum sector is
expected, when it has point-field coordinates of the appropriate regularity,
to produce Wightman distributions; when it has Euclidean reflection-positive
correlators with the required growth, to produce OS data; and when it is
organized by bounded localized observables, to produce a Haag--Kastler type
net.  These are expected properties of a completed local quantum theory in the
corresponding sector.  Comparisons among the frameworks require explicit
hypotheses: charged superselection sectors, gauge-invariant observable
algebras, fermionic field variables in path integrals, curved-background
formulations, and functorial or extended geometric structures may be primary
in one presentation and derived or hidden in another.

The path-integral aspiration can be stated as a gluing problem.  One seeks an
assignment of amplitudes and correlation distributions to a declared category
of backgrounds with sources,
\[
  (M,g,\text{bundles},J,\ldots)
  \longmapsto
  Z[M,g,\text{bundles},J,\ldots],
\]
compatible with cutting and gluing, locality of source variation, reflection
or unitarity, and covariance under background isomorphisms.  A
Kontsevich--Segal type formulation packages this aspiration as a functorial
or multilinear assignment on a category of allowed real or complexified
Riemannian bordisms, with reflection positivity/unitarity represented by
additional structure on the target spaces.  This functorial language is a
valuable target for the geometric and path-integral side of the subject.  In
the opening volumes we use Wightman, OS, and local-net frameworks first
because they give the spectral theory, K\"all\'en--Lehmann representation,
Haag--Ruelle scattering theory, LSZ comparison, and analytic continuation
theorems in the form needed for the later construction.

Constructive results calibrate these frameworks.  Table
\ref{tab:constructive-master-status} in
Chapter~\ref{ch:constructive-status-routes} records representative examples
and open problems.  The completed non-Gaussian constructions are concentrated
in special low-dimensional regimes, such as \(P(\phi)_2\), massive
\(\phi^4_3\), selected two-dimensional scalar--fermion models, massive
Gross--Neveu type models, and two-dimensional Yang--Mills in holonomy or
heat-kernel formulations.  The standard reflection-positive
Ising/\(\lambda\phi^4\) route in \(D\ge4\) is constrained by triviality
theorems, with the four-dimensional marginal case requiring the precise
qualification stated in Table~\ref{tab:constructive-master-status}.  A
four-dimensional UV-complete interacting QFT agreeing with formal
\(\phi^4_4\) perturbation theory to all orders is not ruled out by the
standard positive-lattice triviality theorem as a purely logical matter; it is
a distinct open construction problem.  Four-dimensional Yang--Mills with mass
gap remains the central constructive gauge-theory problem.

\begin{openproblem}[Axiomatic comparison and physical examples]
\label{op:axiomatic-comparison-physical-examples}
Give sharp hypotheses, and prove the resulting comparison theorems, relating
the principal nonperturbative frameworks for local QFT:
\[
\begin{gathered}
  \text{Wightman fields},\qquad
  \text{Osterwalder--Schrader data},\\
  \text{Haag--Kastler local nets},\qquad
  \text{functorial/path-integral data}.
\end{gathered}
\]
Some directions are established under substantial hypotheses: OS-II
reconstruction gives Wightman boundary values from OS-admissible Schwinger
data; Wightman fields give local nets after the relevant self-adjointness,
bounded-functional-calculus, and strong-locality conditions are imposed; and
positive states on local algebras give Hilbert-space representations by GNS.
The converse directions, the precise functorial completion, and the
verification of all required hypotheses in physically central interacting
four-dimensional theories remain open.  The problem has two parts: prove the
comparison maps with minimal usable assumptions, and construct substantial
examples for which the assumptions are actually verified.
\end{openproblem}

\section{Analytic Status of the Basic Objects}

The opening framework deliberately separates several levels of mathematical
status.  This separation is part of the subject, because the same formula may
denote different objects in different constructions.

\begin{definition}[Regulated theory]
  A regulated theory is a finite-dimensional or cutoff-defined construction
  with a specified state space or configuration space, algebra of variables,
  covariance or symmetry action, and expectation functional
  \(\langle\cdot\rangle_\Lambda\), where \(\Lambda\) denotes the regulator
  data.  Examples include a finite lattice, a finite spatial volume with
  momentum cutoff, or a finite-dimensional time-sliced integral.  All
  integrals and operator products in a regulated theory are ordinary objects
  in the regulator-dependent category.
\end{definition}

\begin{definition}[Continuum correlation theory]
  A continuum correlation theory is a hierarchy of distributions, Schwinger
  functions, Wightman functions, or time-ordered distributions obtained as a
  limit of regulated theories or specified directly by axioms.  The topology
  of convergence is part of the datum: distributional convergence of
  correlators, convergence of moments, weak convergence of measures, or
  another declared topology.
\end{definition}

\begin{definition}[Contact terms and local source charts]
\label{def:contact-terms-source-chart}
Let \(M\) be the spacetime manifold and let
\[
  \Delta_n=\bigcup_{1\le i<j\le n}\{(x_1,\ldots,x_n)\in M^n\mid x_i=x_j\}
\]
be the collision locus.  A separated-point \(n\)-point distribution is a
distribution \(T^\circ\in\mathcal D'(M^n\setminus\Delta_n)\), with values in
the appropriate tensor product of operator-label bundles.  A contact-term
extension of \(T^\circ\) is a distribution \(T\in\mathcal D'(M^n)\) whose
restriction to \(M^n\setminus\Delta_n\) equals \(T^\circ\).  A contact term is
a distribution supported on \(\Delta_n\), or on one of its partial collision
diagonals.  Locally near a smooth partial diagonal, with coordinates \(z\) on
the diagonal and normal coordinates \(y\), such a distribution has finite
normal order,
\[
  C
  =
  \sum_{|\alpha|\le N}
  \partial_y^\alpha\delta(y)\,c_\alpha(z),
  \qquad
  c_\alpha\in\mathcal D'(\Delta),
\]
after choosing local coordinates and a density convention.  Thus contact
terms are ordinary distributional data on collision strata.

A local source chart consists of source fields \(J^A\in C_c^\infty(M,E_A)\),
a source functional \(W[J]\) in the chosen category, and distributional
source derivatives
\[
  G_{A_1\cdots A_n}(x_1,\ldots,x_n)
  =
  \left.
  \frac{\delta^n W[J]}
       {\delta J^{A_1}(x_1)\cdots\delta J^{A_n}(x_n)}
  \right|_{J=0}.
\]
Changing the chart by a local functional
\(\int_M\mathcal L(x,j_x^kJ)\dd^Dx\), or by a local source-coordinate change
whose nonlinear part is supported through jets at the same point, changes
these derivatives by finite sums of distributions supported on collision
diagonals.  With the linear source normalization fixed, the restriction of
\(G_{A_1\cdots A_n}\) to \(M^n\setminus\Delta_n\) is unchanged by such
purely local finite choices, while the contact-term extension changes.
\end{definition}

The Wightman and Osterwalder--Schrader frameworks assign distributional
correlators to the selected elementary or named fields.  This assignment gives
the distributions specified by those frameworks, including their declared
domains and positivity properties.  It does not canonically choose composite
fields, time-ordered products, nonlinear source couplings, or the response of
correlators to repeated local source differentiation on collision diagonals.
Those choices are additional local source or operator-product data.  Ward
identities, current algebra, stress-tensor products, anomaly formulae, and
deformations of a theory therefore carry contact-term conventions in addition
to their separated-point correlation functions.

Renormalization is one construction in which these local extension choices
are produced and compared.  The concept is broader: a contact term is
coincident-point/source-response data, while a counterterm is a local term in
a chosen regulated or perturbative action functional.  Perturbative
Epstein--Glaser constructions, perturbative algebraic QFT, microlocal
extension methods, Hollands--Wald local Wick-polynomial and time-ordered
product constructions, and Costello--Gwilliam factorization algebras give
precise comparison frameworks for such extensions under their stated
hypotheses.  They will be used as technical models where their hypotheses
match the problem under discussion, without being treated as a complete
nonperturbative axiom system for all QFTs.

\begin{definition}[Hilbert-space or algebraic completion]
  A Hilbert-space or algebraic completion is a reconstruction of operators,
  states, local algebras, and symmetries from correlation data or from an
  abstract local net.  Positivity is the bridge: Hilbert spaces arise from
  positive inner products, while local observable algebras arise from
  positive states and their GNS representations.
\end{definition}

Perturbation theory has its own status.  A perturbative statement is an
asymptotic expansion in a specified parameter, with coefficients computed from
regulated distributions and then renormalized.  The existence of the
perturbative coefficients and the existence of a nonperturbative continuum
object are distinct assertions.  A perturbative assertion is specified when
the observable being expanded, the expansion parameter, the renormalization
prescription, and the error class or asymptotic meaning have been stated.

These levels of status will be tracked separately in later chapters.  Spectral
measures, scattering states, renormalization-group flows, anomalies,
conformal data, and gauge-invariant observables are different constructions
related by theorems, controlled approximations, or comparison problems.

\section{Fields As Local Distributional Data}

A field presentation assigns localized distributional operators to test
functions. For a test function \(f\in C_c^\infty(\mathbb M^D)\) with
\(\operatorname{supp}f\subset\mathcal O\),
\[
  \widehat\Phi_A(f)
  =
  \int_{\mathbb M^D}\dd^Dx\,f(x)\widehat\Phi_A(x)
\]
is localized in \(\mathcal O\), in the sense that the smeared operator is
affiliated with the local algebra \(\Obs(\mathcal O)\) or belongs to a common
dense invariant field domain associated with that algebra. The precise
operator-domain statement will be made in the field framework of the next
chapter.

The field presentation is used in two constructions. First, it gives concrete
generators or affiliated operators for examples. Second, it gives correlation
functions:
\[
  W_{A_1\cdots A_n}(x_1,\ldots,x_n)
  =
  \bra{\vac}
  \widehat\Phi_{A_1}(x_1)\cdots
  \widehat\Phi_{A_n}(x_n)
  \ket{\vac},
\]
again as distributions. Under suitable positivity, covariance, spectral, and
domain assumptions, such correlation functions can reconstruct the Hilbert
space and fields. Reconstruction theorems will be used later as comparison
results with stated hypotheses.

\section{Additional Constructions And Hypotheses}

Several familiar objects are constructed from the framework after adding
further hypotheses.

\begin{itemize}
  \item A stable one-particle sector is a spectral subspace of the translation
  generators associated with an isolated positive-energy mass hyperboloid
  \(p^2=-m^2\), \(p^0>0\), possibly with finite internal or spin
  multiplicity.
  \item An S-matrix is constructed from incoming and outgoing asymptotic
  multi-particle states and wave operators comparing asymptotic Hilbert spaces
  with the interacting Hilbert space.  This is an additional long-time
  dynamical construction.  Surjectivity of those wave operators onto the
  physical sector is the stronger asymptotic-completeness property and must be
  supplied as a separate theorem or hypothesis.
  \item LSZ reduction is a theorem relating S-matrix elements to residues or
  pole parts of time-ordered correlation functions, after the relevant
  asymptotic states and field-particle overlaps have been identified.
  \item A path-integral presentation is a construction or computational
  scheme for correlation functions after a regulator, contour, measure, or
  perturbative prescription has been specified. The symbol
  \(\int D\phi\,\ee^{\ii S[\phi]}\) receives its meaning from that additional
  specification.
\end{itemize}

This dependency order is part of the construction. Perturbative Feynman graphs
may be introduced as an expansion for Green functions before scattering is
discussed. Diagrams for S-matrix elements acquire that interpretation after
asymptotic states and LSZ reduction have been established.

\section{The First Spectral Question}

Translation invariance permits simultaneous spectral analysis of the
energy-momentum generators \(P^\mu\).  Let \(E_P(\Delta)\) denote the joint
spectral projection of \(P=(P^0,\ldots,P^{D-1})\) associated with a Borel set
\(\Delta\subset\mathbb R^D\).  The joint spectrum
\(\operatorname{Spec}(P)\) is the support of this projection-valued measure.
The spectrum condition is
\[
  \operatorname{Spec}(P)\subset
  \overline V_+
  =
  \{p\in\mathbb R^D:p^0\ge0,\ p^2\le0\}.
\]
For \(\mu^2\ge0\), write \(\Delta_F(x;\mu^2)\) for the free scalar Feynman
distribution with mass parameter \(\mu\):
\[
  \Delta_F(x;\mu^2)
  =
  -\ii\lim_{\epsilon\downarrow0}
  \int\frac{\dd^Dk}{(2\pi)^D}\,
  \frac{\ee^{\ii k\cdot x}}{k^2+\mu^2-\ii\epsilon},
\]
where the limit is taken in the sense of tempered distributions.  This is only
a notation-setting definition at this point; its relation to time ordering and
spectral measures is derived in Chapter~\ref{ch:kallen-lehmann}.
For a scalar Wightman field in the vacuum Hilbert-space framework, positivity
of the Hilbert-space inner product, translation invariance, and the spectrum
condition imply that the two-point function admits a positive spectral
measure on invariant mass.  Let \(\dd\rho(\mu^2)\) denote this positive Borel
measure on \([0,\infty)\).  The time-ordered two-point function has the form
\[
  \bra{\vac}T\{\widehat\phi(x)\widehat\phi(0)\}\ket{\vac}
  =
  \int_{[0,\infty)} \dd\rho(\mu^2)\,\Delta_F(x;\mu^2).
\]
When this measure is absolutely continuous with respect to Lebesgue measure,
one may write
\[
  \dd\rho(\mu^2)=\rho_{\rm ac}(\mu^2)\,\dd\mu^2,
  \qquad
  \rho_{\rm ac}(\mu^2)\ge0
  \quad\text{for almost every }\mu^2 .
\]
This formula will be derived from the stated Hilbert-space assumptions in
Chapter~\ref{ch:kallen-lehmann}.  It is the first result in the text that
relates local two-point data to particle content: isolated atoms of the spectral
measure correspond to stable one-particle states, while continuous spectral
support corresponds to multi-particle and other continuum contributions.

\section{Logical Order Of The Opening Construction}

The opening construction will proceed as follows:
\begin{center}
\begin{tabular}{c}
Hilbert space, symmetry, spectrum, vacuum, locality\\[2pt]
\(\longrightarrow\) local fields and Wightman functions
\(\longrightarrow\) path-integral constructions of Green functions\\[2pt]
\(\longrightarrow\) K\"all\'en--Lehmann spectral representation
\(\longrightarrow\) perturbative Green functions\\[2pt]
\(\longrightarrow\) Haag--Ruelle asymptotic states
\(\longrightarrow\) LSZ
\(\longrightarrow\) perturbative scattering amplitudes.
\end{tabular}
\end{center}
Each arrow indicates an added construction or theorem. The construction will
use each arrow only after its assumptions have been stated.

The first detailed computations follow three dependency bands.  First, finite
and regulated Lagrangian quantum mechanics supplies path integrals,
regularization, perturbation theory for correlation functions, diagrammatics,
and local counterterms.  Second, relativistic scalar fields supply the first
local field examples, Green functions, spectral representations, asymptotic
states, the nonperturbative scattering operator, and LSZ.  Third, spin and
gauge fields require representation theory, fermionic and gauge-field
quantization, gauge fixing, ghosts, BRST cohomology, and applications such as
QED precision observables.  Perturbative Feynman rules for scattering
amplitudes belong after the second band has supplied their interpretation.
