\chapter{Haag--Ruelle Scattering Theory}
\label{ch:haag-ruelle}
\ConventionsLorentzian


\section{Scattering States as Hilbert-Space Limits}

The previous chapters constructed local fields, their spectral content, and
Lorentzian time-ordered Green functions. We now add a new kind of object:
asymptotic particle states. These are vectors in the physical Hilbert space
obtained as large-time limits of localized wave packets. Once these limits are
constructed, the scattering operator is a comparison between two Hilbert-space
identifications: one made in the far past and one made in the far future.

The construction in this chapter is for a massive stable bosonic particle. It
uses a Hilbert space \(\Hilb\), a vacuum vector \(\vac\), local operators, the
Poincare representation, the spectrum condition, and an isolated
one-particle mass shell. These hypotheses are stated explicitly because the
large-time limit is a theorem under hypotheses, not an algebraic manipulation
of Green functions.

This chapter is the introductory construction layer.  It works with scalar
local fields and displays the velocity-support geometry, spectral filtering,
and Cook estimates that make the large-time limits plausible and usable in
physics applications.  The theorem-level algebraic statement, formulated for
vacuum local nets and almost-local operators, is
Theorem~\ref{thm:algebraic-haag-ruelle-scattering} in
Chapter~\ref{ch:volume-iv-haag-ruelle-mathematical-scattering}.  The present
chapter should therefore be read as the point-field model and estimate layer
feeding the later theorem, not as a second independent foundation of
scattering theory.

Let spacetime have dimension \(D\), and write \(d=D-1\). The mostly-plus
metric is
\[
  \eta_{\mu\nu}=\operatorname{diag}(-,+,\ldots,+),
  \qquad
  p^2=\eta_{\mu\nu}p^\mu p^\nu .
\]
Let \(U(a,\Lambda)\) be a strongly continuous unitary representation of the
proper orthochronous Poincare group on \(\Hilb\). Its translation subgroup is
written
\[
  U(a,\mathbf 1)=\ee^{\ii a_\mu P^\mu},
\]
where the commuting self-adjoint generators \(P^\mu\) are the
energy-momentum operators. The spectrum condition is
\[
  \operatorname{Spec}(P)\subset \overline V_+
  =
  \{p\in\mathbb R^D:p^0\ge 0,\ p^2\le 0\}.
\]
The invariant mass operator is
\[
  M^2=-P_\mu P^\mu .
\]

\begin{definition}[Massive one-particle sector]
\label{def:hr-massive-one-particle-sector}
Fix \(m>0\). A massive one-particle sector of mass \(m\) is a closed subspace
\(\Hilb_1\subset\Hilb\) such that:
\begin{enumerate}
  \item \(\Hilb_1\) reduces the Poincare representation;
  \item \(M^2=m^2\) on \(\Hilb_1\);
  \item the spectrum of \(P^\mu\) on \(\Hilb_1\) lies on
  \[
    \Sigma_m^+
    =
    \{(\omega_{\vec p},\vec p):\vec p\in\mathbb R^d\},
    \qquad
    \omega_{\vec p}=\sqrt{\vec p^{\,2}+m^2};
  \]
  \item there is an open interval around \(m^2\) in the spectrum of \(M^2\)
  whose only spectral point is \(m^2\).
\end{enumerate}
Let \(P_1\) denote the orthogonal projection from \(\Hilb\) onto \(\Hilb_1\).
\end{definition}

The last condition is the mass-gap form of stability used in this chapter. It
says that the one-particle shell is separated from the continuum. In a scalar
theory with a lightest stable particle, the picture is:
\[
\begin{tikzpicture}[baseline=-0.55ex,scale=0.94,>=Stealth]
  \draw[->,line width=0.65pt] (-0.2,0)--(7.2,0)
    node[right] {$M^2=-P^2$};
  \draw[line width=0.75pt] (0,0.12)--(0,-0.12);
  \node[below] at (0,-0.12) {vacuum};
  \draw[blue!75!black,line width=1.1pt] (2.15,0.32)--(2.15,-0.32);
  \node[blue!75!black,below] at (2.15,-0.32) {$m^2$};
  \draw[red!75!black,line width=2.2pt] (4.65,0)--(7.0,0);
  \node[red!75!black,below] at (4.65,-0.18) {$M_{\mathrm{thr}}^2$};
  \draw[decorate,decoration={brace,amplitude=5pt},gray!75]
    (2.35,0.55)--(4.45,0.55);
  \node[gray!75!black] at (3.5,0.95) {spectral gap};
\end{tikzpicture}
\]
Here \(M_{\mathrm{thr}}\) is the lowest invariant mass in the continuum channel
under discussion.

For the scalar multiplicity-one sector used below, the abstract Hilbert space
\(\Hilb_1\) is represented concretely as wave functions on the positive mass
shell.  Let
\[
  \dd\mu_m(p)
  =
  \frac{\dd^d\vec p}{(2\pi)^d2\omega_{\vec p}},
  \qquad
  p=(\omega_{\vec p},\vec p).
\]
After fixing the relativistic normalization of one-particle states, choose a
unitary map
\[
  W_1:L^2(\Sigma_m^+,\dd\mu_m)\longrightarrow \Hilb_1
\]
which intertwines translations by
\[
  U(a,\mathbf 1)W_1 f
  =
  W_1\bigl(p\mapsto \ee^{\ii a_\mu p^\mu}f(p)\bigr).
\]
In generalized momentum notation this is
\[
  W_1 f
  =
  \int_{\Sigma_m^+}\dd\mu_m(p)\,f(p)|p\rangle,
  \qquad
  \int_{\Sigma_m^+}\dd\mu_m(p)\,|p\rangle\langle p|
  =
  \mathbf 1_{\Hilb_1}.
\]
The kets in this display are distributional symbols for the spectral
representation; Hilbert vectors are obtained after \(L^2\)-smearing.  With an
internal multiplicity space or nontrivial spin, the scalar model is replaced by
\[
  \Hilb_1\simeq L^2(\Sigma_m^+,\dd\mu_m;\mathfrak k),
\]
where \(\mathfrak k\) carries the little-group/internal representation and the
Poincare action contains the corresponding Wigner rotations.  The present
chapter takes \(\mathfrak k=\mathbb C\).

\begin{remark}[When the isolated shell is a theorem]
\label{rem:haag-ruelle-isolated-shell-existence-scope}
The definition above is an assumption on the Hilbert-space spectrum.  It is
automatic for a free massive scalar field, and it is a proved property in
constructive massive scalar models only when the construction supplies both the
Wightman theory and the relevant mass-gap/one-particle spectral theorem.  In
particular, low-dimensional \(P(\phi)_2\) and massive \(\lambda\phi^4_3\)
models provide the standard constructive setting for such statements in stable
massive sectors.  By contrast, a four-dimensional perturbative
\(\lambda\phi^4\) calculation by itself does not prove the existence of a
UV-complete Hilbert-space theory with an isolated mass shell.  In QED,
long-range soft photons destroy the sharp charged-particle mass hyperboloid
expected by this definition; in QCD, colored quark and gluon shells are absent
from the gauge-invariant observable spectrum.  Stable hadrons, if isolated
below decay thresholds, may satisfy the hypothesis, while resonances and
unstable particles do not.
\end{remark}

\section{Local Operators and One-Particle Wave Packets}

To state the large-time construction without domain distractions, use bounded
local operators. A bounded operator \(A\) is localized in a bounded spacetime
region \(O\) when it belongs to the local algebra \(\Obs(O)\). Locality means
that if \(O_1\) and \(O_2\) are spacelike separated, then
\[
  [A_1,A_2]=0,
  \qquad
  A_i\in\Obs(O_i).
\]
For unbounded point fields, this is the corresponding statement after
smearing and restricting to the common invariant domain on which the products
are defined.

\begin{definition}[Almost-local bounded operator]
\label{def:almost-local-bounded-operator}
Let \(\mathcal O_R\) be a cofinal family of double cones whose radii tend to
infinity; for example, one may take
\[
  \mathcal O_R=\{(x^0,\vec x): |x^0|+|\vec x|<R\},
  \qquad R\ge1 .
\]
A bounded operator \(B\in\mathcal B(\Hilb)\) is \emph{almost local} if there
are bounded operators \(B_R\in\Obs(\mathcal O_R)\) such that, for every integer
\(N\ge1\), there is a constant \(C_N<\infty\) with
\[
  \|B-B_R\|\le C_N R^{-N}
  \qquad (R\ge1).
\]
\end{definition}

The particle must be creatable from the vacuum by local operations. In the
present scalar setting this means that there is a local operator \(A\) such
that
\[
  P_1A\vac\ne0 .
\]

\begin{definition}[Spectrally filtered interpolating creator]
\label{def:hr-spectrally-filtered-interpolating-creator}
Choose a Schwartz test function \(\chi\in\mathcal S(\mathbb R^D)\) whose
Fourier transform is supported in a small neighborhood of the positive mass
shell \(\Sigma_m^+\) and avoids the rest of the energy-momentum spectrum.
Define
\[
  B=A(\chi)
  =
  \int_{\mathbb R^D}\dd^D x\,\chi(x)\,A(x),
  \qquad
  A(x)=U(x)AU(x)^{-1}.
\]
\end{definition}

\begin{proposition}[Filtered local creators are almost local]
\label{prop:hr-filtered-local-creators-almost-local}
If \(A\in\Obs(O)\) is bounded and \(\chi\in\mathcal S(\mathbb R^D)\), then
the filtered operator \(B=A(\chi)\) of
Definition~\ref{def:hr-spectrally-filtered-interpolating-creator} is almost
local.  If the support of \(\widehat\chi\) meets the joint spectrum of
\(P^\mu\) only on the isolated one-particle shell in
Definition~\ref{def:hr-massive-one-particle-sector}, then \(B\vac\) has no
spectral component outside \(\Hilb_1\) in that support window.
\end{proposition}

\begin{proof}
Let \(O\) be contained in a double cone of radius \(R_0\).  Choose smooth
cutoff functions \(\theta_R\) on spacetime, equal to \(1\) on
\(\{x:|x^0|+|\vec x|\le R\}\), supported in
\(\{x:|x^0|+|\vec x|\le 2R\}\), and with derivatives bounded uniformly after
rescaling.  Define
\[
  B_R=\int \dd^Dx\,\theta_R(x)\chi(x)A(x).
\]
The integral is a norm-convergent Bochner integral because
\(\chi\in\mathcal S(\mathbb R^D)\) and \(\|A(x)\|=\|A\|\).  Since
\(\operatorname{supp}\theta_R\) is bounded and \(A(x)\) is localized in
\(O+x\), the operator \(B_R\) belongs to the local algebra of a double cone
whose radius is \(2R+R_0\).  Reindexing the cofinal family of double cones is
irrelevant in the definition of almost locality.

The tail is bounded in operator norm by
\[
  \|B-B_R\|
  \le
  \|A\|\int_{|x^0|+|\vec x|\ge R}\dd^Dx\,|\chi(x)|.
\]
For every \(N\), the Schwartz seminorm
\(\sup_x(1+|x^0|+|\vec x|)^{N+D+1}|\chi(x)|\) is finite, and the last integral
is bounded by \(C_N R^{-N}\).  Hence \(B\) is almost local.

For the spectral statement, write the joint spectral measure of \(P^\mu\) as
\(E(\cdot)\) and use the Fourier convention
\[
  \widehat\chi(p)=\int\dd^Dx\,\chi(x)\ee^{\ii x_\mu p^\mu}.
\]
Because \(U(x)\vac=\vac\),
\[
  B\vac
  =
  \int\dd^Dx\,\chi(x)U(x)A\vac
  =
  \int_{\operatorname{Spec}(P)}\widehat\chi(p)\,\dd E(p)A\vac ,
\]
where the last equality is the joint spectral theorem applied to the vector
\(A\vac\).  Thus \(B\vac\) has spectral support contained in
\(\operatorname{supp}\widehat\chi\cap\operatorname{Spec}(P)\).  If this set is
contained in the isolated mass shell \(\Sigma_m^+\), the corresponding vector
lies in \(E(\Sigma_m^+)\Hilb=\Hilb_1\), and all components outside
\(\Hilb_1\) in the chosen support window vanish.
\end{proof}

Let \(h\in C_c^\infty(\mathbb R^d)\) be a smooth compactly supported momentum
wave packet. Define the positive-energy Klein--Gordon solution
\[
  h_t(\vec x)
  =
  \int\frac{\dd^d\vec p}{(2\pi)^d}\,
  h(\vec p)\,
  \ee^{\ii\omega_{\vec p}t-\ii\vec p\cdot\vec x}.
\]
The associated Haag--Ruelle operator is
\[
  B_t(h)
  =
  \int_{\mathbb R^d}\dd^d\vec x\,
  h_t(\vec x)\,B(t,\vec x),
  \qquad
  B(t,\vec x)=U(t,\vec x)BU(t,\vec x)^{-1}.
\]
The one-particle vector created by this family is
\[
  \psi_h:=P_1B_t(h)\vac .
\]
It is independent of \(t\). The role of the phase in \(h_t\) is precisely to
cancel the free time evolution on the mass shell.

\paragraph{Spectral filter and one-particle time independence.}
Let \(B\) be an almost-local operator whose energy-momentum transfer is
contained in a neighborhood of \(\Sigma_m^+\) disjoint from the rest of
\(\operatorname{Spec}(P)\).  Write the one-particle spectral wavefunction of
\(B\vac\) as
\[
  P_1B\vac
  =
  W_1 b,\qquad
  b\in L^2(\Sigma_m^+,\dd\mu_m).
\]
For \(h\in C_c^\infty(\mathbb R^d)\), the Haag--Ruelle approximant satisfies
\[
  P_1B_t(h)\vac
  =
  W_1(bh),
  \qquad
  P_1\frac{\dd}{\dd t}B_t(h)\vac=0.
\]
In particular, the one-particle vector represented by \(B_t(h)\vac\) is
independent of \(t\).
The verification is the mass-shell phase cancellation: because
\(U(t,\vec x)\vac=\vac\) and \(P_1\) commutes with translations,
the mass-shell component of \(B(t,\vec x)\vac\) is
\[
  P_1B(t,\vec x)\vac
  =
  W_1\!\left(
    p\mapsto
    \ee^{-\ii\omega_{\vec p}t+\ii\vec p\cdot\vec x}b(p)
  \right),
\]
using \(a_\mu p^\mu=-t\omega_{\vec p}+\vec x\cdot\vec p\).  Substitution in
the definition of \(B_t(h)\) gives
\[
\begin{aligned}
P_1B_t(h)\vac
&=
W_1\!\left(
  p\mapsto b(p)
  \int_{\mathbb R^d}\dd^d\vec x\,
  h_t(\vec x)
  \ee^{-\ii\omega_{\vec p}t+\ii\vec p\cdot\vec x}
\right)  \\
&=
W_1\!\left(
  p\mapsto b(p)
  \int\frac{\dd^d\vec q}{(2\pi)^d}\,h(\vec q)
  \ee^{\ii(\omega_{\vec q}-\omega_{\vec p})t}
  \int_{\mathbb R^d}\dd^d\vec x\,
  \ee^{\ii(\vec p-\vec q)\cdot\vec x}
\right)  \\
&=W_1(bh).
\end{aligned}
\]
The last expression is independent of \(t\).  Differentiating the first line
under the smooth compactly supported momentum integral gives
\(P_1\frac{\dd}{\dd t}B_t(h)\vac=0\).

\section{Point-Field Smearing and Propagation}

The bounded-operator formulation is the cleanest theorem statement.  In
applications one often starts from a scalar Wightman field \(\hat\phi(x)\).  The
same construction is then expressed by smearing \(\hat\phi\) against spacetime
test functions and by choosing the Fourier support so that only the isolated
one-particle shell is selected.

\begin{proposition}[Point-field projection to the isolated shell]
\label{prop:hr-point-field-isolated-shell-projection}
Assume that \(\hat\phi\) is a scalar local field whose matrix element between
the vacuum and the one-particle sector is nonzero.  The kets below are
generalized vectors in the rigged-Hilbert-space sense of
Section~\ref{sec:rigged-hilbert-continuous-spectrum}. With relativistically
normalized generalized one-particle kets,
\[
  \langle q|p\rangle
  =
  (2\pi)^d2\omega_{\vec p}\,\delta^{(d)}(\vec q-\vec p),
  \qquad
  \int_{\Sigma_m^+}\dd\mu_m(p)\,|p\rangle\langle p|=\mathbf 1_{\Hilb_1},
\]
where
\[
  \dd\mu_m(p)
  =
  \frac{\dd^d\vec p}{(2\pi)^d2\omega_{\vec p}},
  \qquad
  p=(\omega_{\vec p},\vec p),
\]
the one-particle part of \(\hat\phi(x)\vac\) has the form
\[
  P_1\hat\phi(x)\vac
  =
  Z_\phi^{1/2}
  \int_{\Sigma_m^+}\dd\mu_m(p)\,
  \ee^{-\ii p\cdot x}|p\rangle .
\]
Here \(Z_\phi>0\) is the residue of the two-point function at this particular
field's one-particle pole.  All remaining spectral components of
\(\hat\phi(x)\vac\) have invariant mass separated from \(m\) by the assumed
gap.

Let
\[
  f(x)=\int\frac{\dd^D k}{(2\pi)^D}\,\tilde f(k)\,\ee^{\ii k\cdot x},
  \qquad
  \hat\phi_f=\int\dd^D x\,f(x)\hat\phi(x).
\]
If \(\tilde f\) is supported in a small neighborhood of \(\Sigma_m^+\) and away
from the rest of the energy-momentum spectrum, then
\[
  \hat\phi_f\vac
  =
  Z_\phi^{1/2}
  \int_{\Sigma_m^+}\dd\mu_m(p)\,
  \tilde f(p)|p\rangle ,
\]
with no continuum contribution.
\end{proposition}

\begin{proof}
The point-field expression is a distributional statement.  The actual Hilbert
vectors are obtained after smearing in \(x\).  In the scalar multiplicity-one
sector, define the one-particle form-factor distribution by
\[
  F_x(p)=\langle p|\hat\phi(x)\vac\rangle .
\]
Translation covariance and \(U(x)\vac=\vac\) give
\[
  F_x(p)
  =
  \langle p|U(x)\hat\phi(0)\vac\rangle
  =
  \ee^{-\ii p\cdot x}F_0(p),
\]
because \(U(x)|p\rangle=\ee^{\ii p\cdot x}|p\rangle\).  Scalar Lorentz
covariance gives \(F_0(\Lambda p)=F_0(p)\) on the positive mass shell.  Since
\(\Sigma_m^+\) is a single \(SO^+(1,d)\)-orbit and the multiplicity space is
\(\mathbb C\), \(F_0\) is a constant distribution on the shell.  After fixing
the phase of the interpolating field, write this constant as
\(Z_\phi^{1/2}>0\); equivalently \(Z_\phi=|F_0|^2\) is the coefficient of the
one-particle delta mass in the K\"all\'en--Lehmann measure of this field.

Therefore, in the rigged-Hilbert notation,
\[
  P_1\hat\phi(x)\vac
  =
  Z_\phi^{1/2}
  \int_{\Sigma_m^+}\dd\mu_m(p)\,\ee^{-\ii p\cdot x}|p\rangle .
\]
For a Schwartz smearing function with
\[
  f(x)=\int\frac{\dd^D k}{(2\pi)^D}\tilde f(k)\ee^{\ii k\cdot x},
\]
pairing this distribution with \(f\) gives
\[
  P_1\hat\phi_f\vac
  =
  Z_\phi^{1/2}
  \int_{\Sigma_m^+}\dd\mu_m(p)
  \left(\int\dd^Dx\,f(x)\ee^{-\ii p\cdot x}\right)|p\rangle
  =
  Z_\phi^{1/2}
  \int_{\Sigma_m^+}\dd\mu_m(p)\,\tilde f(p)|p\rangle .
\]
The remaining spectral part of the smeared vector \(\hat\phi_f\vac\) is
obtained by applying the joint spectral measure of \(P^\mu\) to the part of
the field-created distribution supported away from the isolated shell.  If
\(\operatorname{supp}\tilde f\) is contained in the chosen isolated-shell
neighborhood and avoids the rest of the spectrum, that spectral projection is
zero.  The displayed mass-shell integral is then the full vector
\(\hat\phi_f\vac\).
\end{proof}

A Gaussian packet centered near a spacetime point \((t_*,\vec x_*)\) and a
momentum \((E_*,\vec P_*)\) may be used as an approximation,
\[
  f(x)
  =
  \exp\left[
    -\frac{(x^0-t_*)^2+|\vec x-\vec x_*|^2}{L^2}
    +\ii\vec P_*\cdot(\vec x-\vec x_*)
    -\ii E_*(x^0-t_*)
  \right],
  \qquad E_*^2=\vec P_*^{\,2}+m^2 .
\]
For exact spectral separation it is better to choose \(\tilde f\) smooth and
compactly supported in momentum space; the spacetime function is then rapidly
decreasing rather than compactly supported.  This is the analytic setting used
by Haag--Ruelle theory.

Define the translated family \(f^{(T)}\) by
\[
  \tilde f^{(T)}(k)
  =
  \ee^{\ii(k^0-\omega_{\vec k})T}\tilde f(k),
  \qquad
  \omega_{\vec k}=\sqrt{\vec k^{\,2}+m^2}.
\]
On the positive mass shell the extra phase is equal to \(1\).  Consequently
\[
  P_1\hat\phi_{f^{(T)}}\vac=P_1\hat\phi_f\vac
\]
for all \(T\), while the spacetime profile of the smearing function moves.  In
position space,
\[
\begin{aligned}
  f^{(T)}(x^0,\vec x)
  &=
  \int\frac{\dd^Dk}{(2\pi)^D}\,
  \ee^{\ii k\cdot x+\ii(k^0-\omega_{\vec k})T}\tilde f(k)
  \\
  &=
  \int\dd^d\vec y\,K(\vec x-\vec y;T)\,
  f(x^0-T,\vec y),
\end{aligned}
\]
where
\[
  K(\vec x;T)
  =
  \int\frac{\dd^d\vec k}{(2\pi)^d}\,
  \ee^{\ii\vec k\cdot\vec x-\ii\omega_{\vec k}T}.
\]
The kernel satisfies the Klein--Gordon equation in the variables
\((T,\vec x)\),
\[
  \left(\partial_T^2-\Delta_{\vec x}+m^2\right)K(\vec x;T)=0,
  \qquad
  K(\vec x;0)=\delta^{(d)}(\vec x)
\]
in the distributional sense.  If \(\tilde f\) is concentrated near
\(\vec P_*\), stationary phase shows that the packet center moves as
\[
  (t_*,\vec x_*)
  \longmapsto
  (t_*+T,\vec x_*+\vec v_*T),
  \qquad
  \vec v_*=\nabla_{\vec p}\omega_{\vec p}\big|_{\vec p=\vec P_*}
  =\frac{\vec P_*}{E_*}.
\]
Thus the same one-particle vector may be represented by smearing a local field
over different large spacetime regions.  The equality is only after projection
to \(\Hilb_1\); the parts outside the isolated shell are removed by the spectral
support condition.

The velocity support of \(h\) is
\[
  V(h)
  =
  \left\{
    \frac{\vec p}{\omega_{\vec p}}:
    \vec p\in\operatorname{supp}h
  \right\}
  \subset\mathbb R^d .
\]
Stationary phase gives the spacetime localization of \(h_t\): for large
\(|t|\), the wave packet is concentrated near rays
\[
  \vec x=t\,\frac{\vec p}{\omega_{\vec p}},
  \qquad
  \vec p\in\operatorname{supp}h .
\]
Thus two wave packets with disjoint velocity supports become spacelike
separated as \(|t|\to\infty\).

\begin{center}
\begin{tikzpicture}[scale=0.94,>=Stealth]
  \draw[->,line width=0.7pt] (-3.35,-2.0)--(-3.35,2.45)
    node[above] {$x^0$};
  \draw[->,line width=0.7pt] (-3.35,-2.0)--(3.5,-2.0)
    node[right] {$\vec x$};
  \draw[gray!55,dashed] (-3.0,-1.55)--(2.75,1.85);
  \draw[gray!55,dashed] (-1.15,-1.55)--(3.15,1.45);
  \fill[blue!10,draw=blue!75!black,line width=1.0pt]
    (-1.85,-1.1) ellipse (0.58 and 0.34);
  \fill[green!12,draw=green!50!black,line width=1.0pt]
    (-0.6,-1.1) ellipse (0.58 and 0.34);
  \fill[blue!10,draw=blue!75!black,line width=1.0pt]
    (-0.35,0.65) ellipse (0.55 and 0.34);
  \fill[green!12,draw=green!50!black,line width=1.0pt]
    (1.75,0.35) ellipse (0.55 and 0.34);
  \draw[->,blue!75!black,very thick] (-1.65,-0.78)--(-0.62,0.35);
  \draw[->,green!50!black,very thick] (-0.35,-0.78)--(1.38,0.13);
  \node[blue!75!black] at (-1.15,1.03) {$V(h_1)$};
  \node[green!50!black] at (2.25,0.78) {$V(h_2)$};
  \draw[<->,purple!80!black,very thick] (0.22,0.62)--(1.18,0.48);
  \node[purple!80!black,align=center] at (0.73,1.18)
    {spacelike\\separation};
\end{tikzpicture}
\end{center}

\section{Velocity Support Separation}

The phrase ``disjoint velocity supports'' has a precise geometric content.
Let \(K_i=\operatorname{supp}h_i\) and set
\[
  V_i=\left\{\vec p/\omega_{\vec p}:\vec p\in K_i\right\}.
\]
If \(V_i\cap V_j=\varnothing\), compactness gives a number
\[
  \delta_{ij}:=\operatorname{dist}(V_i,V_j)>0 .
\]
Choose \(0<\epsilon<\delta_{ij}/4\) and define the equal-time tubes
\[
  \Gamma_i(t,\epsilon)
  =
  \{\vec x\in\mathbb R^d:\operatorname{dist}(\vec x/t,V_i)<\epsilon\}
  \qquad (t\ne0).
\]
For \(\vec x\in \Gamma_i(t,\epsilon)\) and
\(\vec y\in\Gamma_j(t,\epsilon)\),
\[
  |\vec x-\vec y|
  \ge |t|(\delta_{ij}-2\epsilon)
  \ge \frac12|t|\delta_{ij}.
\]
The spacetime points \((t,\vec x)\) and \((t,\vec y)\) are therefore
spacelike separated with a separation growing linearly in \(|t|\).

The part of \(h_{i,t}\) outside \(\Gamma_i(t,\epsilon)\) is rapidly
decreasing.  Indeed, away from the stationary set
\(\vec x/t=\vec p/\omega_{\vec p}\), the phase
\[
  \omega_{\vec p}t-\vec p\cdot\vec x
\]
has no stationary point on \(K_i\), and repeated integration by parts in
\(\vec p\) gives a bound \(O(|t|^{-N})\) for every \(N\).  Thus the
Haag--Ruelle operators can be replaced, up to rapidly decreasing errors, by
almost-local operators supported in mutually spacelike tubes.  Locality then
turns geometric separation into the commutator estimate used in the
existence proof.

\begin{proposition}[Velocity-tube separation and off-tube decay]
\label{prop:hr-velocity-tube-separation}
Let \(h\in C_c^\infty(\mathbb R^d)\), and let
\[
  V(h)=\{\nabla_{\vec p}\omega_{\vec p}:\vec p\in\operatorname{supp}h\}.
\]
For every \(\epsilon>0\) and every \(N\ge1\), there is \(C_{N,\epsilon}\) such
that
\[
  \sup_{\operatorname{dist}(\vec x/t,V(h))\ge\epsilon}
  |h_t(\vec x)|
  \le
  C_{N,\epsilon}(1+|t|)^{-N},
  \qquad |t|\ge1 .
\]
If \(h_i,h_j\) have compact velocity supports with
\(\delta_{ij}=\operatorname{dist}(V(h_i),V(h_j))>0\), then for
\(0<\epsilon<\delta_{ij}/4\) and
\(\vec x\in\Gamma_i(t,\epsilon)\), \(\vec y\in\Gamma_j(t,\epsilon)\),
\[
  |\vec x-\vec y|\ge \frac12\delta_{ij}|t|.
\]
Thus the equal-time spacetime points \((t,\vec x)\) and \((t,\vec y)\) are
spacelike separated by a distance growing linearly in \(|t|\).
\end{proposition}

\begin{proof}
Write
\[
  h_t(\vec x)
  =
  \int\frac{\dd^d\vec p}{(2\pi)^d}
  h(\vec p)\,
  \ee^{\ii t\Phi_{\vec x/t}(\vec p)},
  \qquad
  \Phi_{\vec u}(\vec p)=\omega_{\vec p}-\vec p\cdot\vec u .
\]
On the compact set \(\operatorname{supp}h\), the gradient is
\[
  \nabla_{\vec p}\Phi_{\vec u}(\vec p)
  =
  \frac{\vec p}{\omega_{\vec p}}-\vec u .
\]
If \(\operatorname{dist}(\vec u,V(h))\ge\epsilon\), then
\(|\nabla_{\vec p}\Phi_{\vec u}|\ge\epsilon\) on \(\operatorname{supp}h\).
Apply repeatedly the adjoint of
\[
  L_{\vec u}
  =
  \frac{1}{\ii t}
  \frac{\nabla_{\vec p}\Phi_{\vec u}\cdot\nabla_{\vec p}}
       {|\nabla_{\vec p}\Phi_{\vec u}|^2},
  \qquad
  L_{\vec u}\ee^{\ii t\Phi_{\vec u}}=\ee^{\ii t\Phi_{\vec u}} .
\]
Derivatives of \(\Phi_{\vec u}\) are uniformly bounded for \(\vec p\) in the
fixed compact support and for the relevant \(\vec u\).  Each integration by
parts gives a factor \(|t|^{-1}\), proving the displayed off-tube estimate.

For the geometric separation, choose
\(\vec v_i\in V(h_i)\) and \(\vec v_j\in V(h_j)\) with
\[
  |\vec x/t-\vec v_i|<\epsilon,\qquad
  |\vec y/t-\vec v_j|<\epsilon .
\]
Then
\[
  |\vec x-\vec y|
  =
  |t|\,|\vec x/t-\vec y/t|
  \ge
  |t|(\delta_{ij}-2\epsilon)
  \ge
  \frac12\delta_{ij}|t|.
\]
Equal-time separation is spacelike when \(\vec x\ne\vec y\), since the
Minkowski squared separation is \(|\vec x-\vec y|^2\).
\end{proof}

\section{Multi-Particle Limits}

Let \(B_1,\ldots,B_n\) be almost-local operators obtained as above, and let
\(h_1,\ldots,h_n\in C_c^\infty(\mathbb R^d)\) have pairwise disjoint velocity
supports. Define the time-dependent vector
\[
  \Psi_t(h_1,\ldots,h_n)
  =
  B_{1,t}(h_1)\cdots B_{n,t}(h_n)\vac .
\]

\begin{proposition}[Cook estimate form of the Haag--Ruelle limits]
\label{prop:haag-ruelle-cook-limits}
Assume the locality, spectrum, isolated-shell, and regularity hypotheses
stated above.  Suppose, in addition, that the Haag--Ruelle product
\(\Psi_t(h_1,\ldots,h_n)\) is strongly differentiable on the large-time tails
and that there are integrable functions \(R_\pm\) such that
\[
  \left\|\frac{\dd}{\dd t}\Psi_t(h_1,\ldots,h_n)\right\|
  \le R_+(t)\quad (t\ge T_0),
  \qquad
  \left\|\frac{\dd}{\dd t}\Psi_t(h_1,\ldots,h_n)\right\|
  \le R_-(t)\quad (t\le -T_0).
\]
Suppose also that the recursive scalar-product contraction estimates hold:
large-time scalar products of velocity-separated products are obtained by
pairing one-particle vectors, up to errors that vanish on the large-time
tails.  Finally, suppose the same Cook and contraction estimates hold after
replacing one creator by the difference of two creators with the same
one-particle vector.  Then the limits
\[
  \Psi^{\mathrm{out}}(h_1,\ldots,h_n)
  =
  \lim_{t\to+\infty}\Psi_t(h_1,\ldots,h_n),
\]
\[
  \Psi^{\mathrm{in}}(h_1,\ldots,h_n)
  =
  \lim_{t\to-\infty}\Psi_t(h_1,\ldots,h_n)
\]
exist in the norm topology of \(\Hilb\). The limits depend only on the
one-particle vectors
\[
  \psi_j=P_1B_{j,t}(h_j)\vac\in\Hilb_1,
  \qquad j=1,\ldots,n,
\]
and not on the auxiliary operators \(B_j\) used to interpolate them.
\end{proposition}

\begin{proof}
For \(t'>t\ge T_0\),
\[
  \|\Psi_{t'}-\Psi_t\|
  \le
  \int_t^{t'}
  \left\|\frac{\dd}{\dd s}\Psi_s\right\|\,\dd s
  \le
  \int_t^{t'}R_+(s)\,\dd s .
\]
Since \(R_+\in L^1([T_0,\infty))\), the right side tends to zero when
\(t,t'\to+\infty\).  Hence \(\Psi_t\) is Cauchy in the Hilbert norm and has
an outgoing limit.  The incoming proof is identical with \(R_-\) on the
negative tail.

If \(B_j\) and \(B'_j\) have the same one-particle vector, multilinearity
writes the difference of the two \(n\)-particle approximants as a finite sum
in which one factor is \(D_j=B_j-B'_j\).  For that factor
\[
  P_1D_{j,t}(h_j)\vac=0 .
\]
The assumed Cook estimate for such zero-one-particle replacements implies
that each difference term has a norm limit.  Its limiting Fock inner product
with all velocity-separated Haag--Ruelle vectors is zero, because the
recursive contraction of asymptotic scalar products contains no one-particle
contraction for the \(D_j\) entry.  Density of the velocity-separated
algebraic Fock domain then gives that this limit vector is zero.  Thus the
large-time vector depends only on \(\psi_j=P_1B_{j,t}(h_j)\vac\).
\end{proof}

The proposition has two analytic inputs. First, disjoint velocity supports imply
that the effective localization regions of \(B_{j,t}(h_j)\) separate
spacelike at large \(|t|\). Locality then gives rapid decay of commutators
\[
  [B_{i,t}(h_i),B_{j,t}(h_j)]
\]
for \(i\ne j\). Second, spectral isolation of the mass shell controls the part
of \(B_{j,t}(h_j)\vac\) orthogonal to \(\Hilb_1\). In Cook's method one
estimates \(\frac{\dd}{\dd t}\Psi_t\); the commutator terms and the
orthogonal spectral-error terms are integrable in \(t\), hence
\(\Psi_t\) is a Cauchy family at \(+\infty\) and at \(-\infty\).

For \(n=2\), the essential estimate is visible from the norm difference
\[
  \|\Psi_{t'}-\Psi_t\|^2
  =
  \langle\Psi_{t'}-\Psi_t,\Psi_{t'}-\Psi_t\rangle .
\]
The scalar products entering this expression are vacuum correlation functions
of four almost-local operators. At large positive or negative times, the two
velocity clusters are separated by spacelike distances growing linearly in
\(|t|\). The cluster property and locality reduce the large-time scalar
product to the product of the corresponding one-particle scalar products.

In the point-field notation above, for two packets with disjoint velocity
supports one may write
\[
  \Psi_T=\hat\phi_{f_1^{(T)}}\hat\phi_{f_2^{(T)}}\vac .
\]
For fixed \(\Delta T\),
\[
\begin{aligned}
  \langle\Psi_T,\Psi_{T+\Delta T}\rangle
  =
  \bra{\vac}
  \hat\phi_{f_2^{(T)}}^*
  \hat\phi_{f_1^{(T)}}^*
  \hat\phi_{f_1^{(T+\Delta T)}}
  \hat\phi_{f_2^{(T+\Delta T)}}
  \ket{\vac} .
\end{aligned}
\]
As \(T\to+\infty\) or \(T\to-\infty\), the two spacetime clusters separate
spacelike.  The vacuum cluster property gives
\[
  \langle\Psi_T,\Psi_{T+\Delta T}\rangle
  -
  \langle\hat\phi_{f_1}\vac,\hat\phi_{f_1}\vac\rangle
  \langle\hat\phi_{f_2}\vac,\hat\phi_{f_2}\vac\rangle
  \longrightarrow 0,
\]
and the same limit is obtained for \(\Delta T=0\).  Therefore
\[
  \|\Psi_{T+\Delta T}-\Psi_T\|\longrightarrow0 .
\]
Under the smooth spectral separation hypotheses of Haag--Ruelle theory, the
error decreases faster than any inverse power of \(|T|\).  This explicit
two-packet calculation is the concrete model for the general Cook estimate.

\section{The Estimate Behind the Limit}

The preceding proposition is a norm-convergence statement.  Its proof is
governed by an estimate of the following form.  Let \(B_{i,t}(h_i)\) and
\(B_{j,t}(h_j)\) be Haag--Ruelle operators with disjoint velocity supports.
For every \(N\ge1\), almost locality and stationary phase give constants
\(C_N\) such that
\begin{equation}
  \left\|
  [B_{i,t}(h_i),B_{j,t}(h_j)]
  \right\|
  \le
  C_N(1+|t|)^{-N}
  \label{eq:haag-ruelle-commutator-decay}
\end{equation}
for sufficiently large \(|t|\).  The estimate says more than that the
operators eventually commute: it gives an integrable error.

The time derivative of a single approximant has no one-particle component:
\[
  P_1\frac{\dd}{\dd t}B_{j,t}(h_j)\vac=0 .
\]
The remaining spectral support lies away from the isolated mass shell.  By
choosing the energy-momentum support of the smearing function \(\chi\) inside
the mass-gap neighborhood, that remainder is controlled by integration by
parts in the energy variable.  In the multi-particle product,
\(\frac{\dd}{\dd t}\Psi_t\) is a sum of terms in which either a one-particle
remainder appears or two factors must be commuted past each other.  The
one-particle remainders are integrable by the spectral gap, and the commutator
terms are integrable by \eqref{eq:haag-ruelle-commutator-decay}.  Therefore
\[
  \int_T^\infty
  \left\|\frac{\dd}{\dd t}\Psi_t\right\|\,\dd t
  \longrightarrow0
  \qquad (T\to\infty),
\]
which is Cook's Cauchy criterion for the outgoing limit.  The incoming limit
is the same argument with \(t\to-\infty\).

This estimate is the point at which locality, spectrum, and particle
stability meet.  Without locality the commutator decay is unavailable; without
the spectral gap the one-particle remainder need not be integrable; without
positive energy the large-time stationary-phase localization has no stable
direction of propagation.

\section{Fock Structure and Wave Operators}

\begin{definition}[Scalar asymptotic Fock space]
\label{def:hr-scalar-asymptotic-fock-space}
Let
\[
  \mathcal F_s(\Hilb_1)
  =
  \bigoplus_{n=0}^\infty \Hilb_1^{\odot n}
\]
be the bosonic Fock space over \(\Hilb_1\), where
\(\Hilb_1^{\odot n}\) denotes the symmetric \(n\)-fold tensor product and
\(\Hilb_1^{\odot0}=\mathbb C\Omega_F\). The vector \(\Omega_F\) is the Fock
vacuum.
\end{definition}

The construction is intrinsic: it is the Hilbert completion of the algebraic
finite symmetric tensor algebra over \(\Hilb_1\).  The mass-shell model above
turns this intrinsic object into the familiar wave-function representation
\[
  \mathcal F_s(\Hilb_1)
  \simeq
  \mathbb C\Omega_F
  \oplus
  \bigoplus_{n\ge1}
  L^2_{\mathrm{sym}}
  \bigl((\Sigma_m^+)^n,\dd\mu_m(p_1)\cdots\dd\mu_m(p_n)\bigr).
\]
Under this identification a decomposable \(n\)-particle vector is represented
by
\[
  (f_1\odot\cdots\odot f_n)(p_1,\ldots,p_n)
  =
  \frac{1}{\sqrt{n!}}
  \sum_{\sigma\in S_n}
  \prod_{j=1}^n f_j(p_{\sigma(j)}),
\]
which gives
\[
  \langle f_1\odot\cdots\odot f_n,
  g_1\odot\cdots\odot g_n\rangle
  =
  \sum_{\sigma\in S_n}
  \prod_{j=1}^n
  \langle f_j,g_{\sigma(j)}\rangle_{L^2(\Sigma_m^+,\dd\mu_m)} .
\]
This is the Fock-space inner product that the Haag--Ruelle limits reproduce
inside the physical Hilbert space.

\begin{remark}[Statistics of the asymptotic Fock space]
The notation \(\mathcal F_s(\Hilb_1)\) is appropriate here because the
one-particle sector under discussion is the scalar bosonic sector constructed
earlier.  For several stable species the asymptotic Fock space is the graded
tensor product determined by their statistics: integer-spin bosonic species
contribute symmetric factors, while half-integer-spin fermionic species
contribute antisymmetric factors.  The spin-statistics theorem gives this
assignment under the Wightman hypotheses stated in
Chapter~\ref{ch:spinor-fields-fermionic-path-integrals}; the present
Haag--Ruelle construction uses the scalar instance.
\end{remark}

\begin{definition}[Haag--Ruelle wave operators on the separated algebraic domain]
\label{def:hr-wave-operators-separated-domain}
The asymptotic identifications from the free Fock space
\(\mathcal F_s(\Hilb_1)\) to the physical Hilbert space are called the
incoming and outgoing Haag--Ruelle wave operators.  On the dense algebraic
domain generated by velocity-separated decomposable vectors, each wave
operator assigns to an abstract collection of asymptotic one-particle wave
packets the physical Hilbert-space vector obtained as a large-time
Haag--Ruelle limit.
\end{definition}

The topology in this statement is pointwise Hilbert-norm convergence on the
asymptotic Fock domain.  Choose, for a dense algebraic subspace
\(\mathcal F_{s,{\rm sep}}^{\rm alg}(\Hilb_1)\subset\mathcal F_s(\Hilb_1)\),
Haag--Ruelle creators and velocity-separated wave packets representing each
decomposable vector \(F\).  The time-dependent identification map
\(\Omega_{{\rm out},t}^{\rm HR}\) is defined on that vector by the corresponding
product of Haag--Ruelle operators applied to the vacuum.  The Cook-limit
statement says
\begin{equation}
  \lim_{t\to+\infty}
  \left\|
    \Omega_{{\rm out},t}^{\rm HR}F-\Omega_{\rm out}F
  \right\|_{\Hilb}
  =0,
  \qquad
  F\in\mathcal F_{s,{\rm sep}}^{\rm alg}(\Hilb_1),
  \label{eq:haag-ruelle-strong-topology-out}
\end{equation}
and analogously for \(t\to-\infty\) and \(\Omega_{\rm in}\).  Packaged as
maps from the asymptotic Fock domain to \(\Hilb\), this is strong operator
convergence on that domain: convergence holds after applying the approximating
map to each fixed vector \(F\).  The weak convergence of all matrix elements
\(\langle\eta,\Omega_{{\rm out},t}^{\rm HR}F\rangle\) follows immediately from
\eqref{eq:haag-ruelle-strong-topology-out}, while the converse weak statement
does not by itself give the isometric wave operator.  Operator-norm
convergence,
\[
  \sup_{\|F\|_{\mathcal F_s}=1}
  \left\|
    \Omega_{{\rm out},t}^{\rm HR}F-\Omega_{\rm out}F
  \right\|_{\Hilb}\longrightarrow0,
\]
is not asserted in Haag--Ruelle theory.  The isometric extension of
\(\Omega_{\rm out/in}\) to all of \(\mathcal F_s(\Hilb_1)\) is obtained after
the Fock inner product has been reproduced on the dense algebraic domain.

\begin{proposition}[Definition and isometry of the scalar Haag--Ruelle maps]
\label{prop:hr-scalar-wave-operator-isometry}
Under the Cook and scalar-product contraction hypotheses of
Proposition~\ref{prop:haag-ruelle-cook-limits}, the following assignments on
velocity-separated decomposable vectors are well-defined:
\[
  \Omega_{\mathrm{out}}
  (\psi_1\odot\cdots\odot\psi_n)
  =
  \Psi^{\mathrm{out}}(h_1,\ldots,h_n),
\]
\[
\Omega_{\mathrm{in}}
  (\psi_1\odot\cdots\odot\psi_n)
  =
  \Psi^{\mathrm{in}}(h_1,\ldots,h_n),
\]
where \(\psi_j=P_1B_{j,t}(h_j)\vac\).  The Cook-limit statement above makes these
definitions independent of the choice of interpolating operators. By
linearity and continuity they extend from such vectors to all of
\(\mathcal F_s(\Hilb_1)\).

The Fock inner product is reproduced by the Hilbert-space inner product of
asymptotic states:
\[
\begin{aligned}
&\left\langle
  \Psi^{\mathrm{out}}(\psi_1,\ldots,\psi_n),
  \Psi^{\mathrm{out}}(\varphi_1,\ldots,\varphi_m)
  \right\rangle_{\Hilb}
\\
&\qquad =
  \delta_{nm}
  \sum_{\sigma\in S_n}
  \prod_{j=1}^n
  \langle\psi_j,\varphi_{\sigma(j)}\rangle_{\Hilb_1},
\end{aligned}
\]
and the same formula holds for incoming states. Therefore
\[
  \Omega_{\mathrm{out}},
  \Omega_{\mathrm{in}}
  :
  \mathcal F_s(\Hilb_1)\longrightarrow \Hilb
\]
are isometric wave operators. Their ranges are the outgoing and incoming
scattering subspaces,
\[
  \Hilb_{\mathrm{out}}=\operatorname{Ran}\Omega_{\mathrm{out}},
  \qquad
  \Hilb_{\mathrm{in}}=\operatorname{Ran}\Omega_{\mathrm{in}} .
\]
\end{proposition}

\begin{proof}
Suppose two choices of almost-local creators and wave packets give the same
one-particle vector \(\psi_j\).  The Cook-limit hypothesis gives large-time
limits for both choices, and the independence part of
Proposition~\ref{prop:haag-ruelle-cook-limits} says that the difference of
the two limiting vectors has Hilbert norm zero.  Hence the assignment depends
only on \(\psi_1,\ldots,\psi_n\), not on the interpolating coordinates used to
produce them.

On the velocity-separated algebraic domain, the scalar-product contraction
estimate of Proposition~\ref{prop:haag-ruelle-cook-limits} gives the
permanent formula displayed above.  That formula is precisely the inner
product of the symmetric Fock tensor
\(\psi_1\odot\cdots\odot\psi_n\) with
\(\varphi_1\odot\cdots\odot\varphi_m\).  Thus the map preserves inner
products on a dense subspace of \(\mathcal F_s(\Hilb_1)\).  A norm-preserving
linear map on a dense subspace extends uniquely to an isometry on the Hilbert
completion, and the range of an isometry is closed because a Cauchy sequence
of image vectors has a Cauchy preimage sequence.
\end{proof}

The Poincare group acts on \(\mathcal F_s(\Hilb_1)\) by the free Fock
representation \(U_F\) induced from its action on \(\Hilb_1\). The
Haag--Ruelle wave operators intertwine the physical and free actions:
\[
  U(a,\Lambda)\Omega_{\mathrm{out/in}}
  =
  \Omega_{\mathrm{out/in}}U_F(a,\Lambda).
\]
This equation states that an asymptotic \(n\)-particle state transforms as
the corresponding symmetrized tensor product of one-particle Wigner states.

Equivalently, the wave operators may be described by generalized momentum
kernels.
For bosons,
\[
\begin{aligned}
&{}^{\mathrm{out/in}}\langle
q_1,\ldots,q_m|p_1,\ldots,p_n\rangle^{\mathrm{out/in}}
\\
&\qquad =
\delta_{mn}\sum_{\sigma\in S_n}
\prod_{j=1}^n
(2\pi)^d2\omega_{\vec p_j}\,
\delta^{(d)}(\vec q_j-\vec p_{\sigma(j)}).
\end{aligned}
\]
For smooth wave packets,
\[
  \Psi^{\mathrm{out/in}}(f_1,\ldots,f_n)
  =
  Z_\phi^{n/2}
  \int\prod_{j=1}^n\dd\mu_m(p_j)\,
  \prod_{j=1}^n\tilde f_j(p_j)\,
  |p_1,\ldots,p_n\rangle^{\mathrm{out/in}},
\]
whenever the \(f_j\) are chosen with separated velocity supports.  This is the
precise form of the statement that either the incoming or the outgoing
asymptotic basis labels states by free particle momenta, while the comparison
between the two bases is the scattering matrix.

\begin{center}
\begin{tikzcd}[column sep=large,row sep=large]
  \mathcal F_s(\Hilb_1)
    \arrow[r,"\Omega_{\mathrm{in}}"]
    \arrow[d,"S"']
  &
  \Hilb_{\mathrm{in}}\subset\Hilb
  \\
  \mathcal F_s(\Hilb_1)
    \arrow[r,"\Omega_{\mathrm{out}}"']
  &
  \Hilb_{\mathrm{out}}\subset\Hilb
    \arrow[u,dashed,leftrightarrow]
\end{tikzcd}
\end{center}

\section{The Scattering Operator}

\begin{definition}[Scattering operator in a massive Haag--Ruelle sector]
\label{def:hr-scattering-operator-massive-sector}
If
\[
  \Hilb_{\mathrm{in}}=\Hilb_{\mathrm{out}},
\]
then incoming scattering states evolve into outgoing scattering states in the
sector under consideration. The scattering operator on the asymptotic Fock
space is
\[
  S
  =
  \Omega_{\mathrm{out}}^*\Omega_{\mathrm{in}}
  :
  \mathcal F_s(\Hilb_1)\longrightarrow \mathcal F_s(\Hilb_1).
\]
\end{definition}

\paragraph{Unitarity and matrix elements of \(S\).}
When the incoming and outgoing ranges coincide, the operator
\[
  S=\Omega_{\mathrm{out}}^*\Omega_{\mathrm{in}}
\]
is unitary.  The matrix
element between an incoming \(n\)-particle wave packet
\(\psi_1\odot\cdots\odot\psi_n\) and an outgoing \(m\)-particle wave packet
\(\varphi_1\odot\cdots\odot\varphi_m\) is
\[
\begin{aligned}
&\left\langle
  \varphi_1\odot\cdots\odot\varphi_m,\,
  S(\psi_1\odot\cdots\odot\psi_n)
  \right\rangle_{\mathcal F_s(\Hilb_1)}
\\
&\qquad =
  \left\langle
  \Omega_{\mathrm{out}}
  (\varphi_1\odot\cdots\odot\varphi_m),\,
  \Omega_{\mathrm{in}}
  (\psi_1\odot\cdots\odot\psi_n)
  \right\rangle_{\Hilb}.
\end{aligned}
\]
This equation is the nonperturbative definition of the scattering matrix
elements in the massive sector. The right hand side is an inner product in the
physical Hilbert space; the left hand side records the same number in the
free asymptotic particle basis.
Let \(\mathcal R=\operatorname{Ran}\Omega_{\mathrm{in}}
=\operatorname{Ran}\Omega_{\mathrm{out}}\).  Because the wave operators are
isometries from \(\mathcal F_s(\Hilb_1)\) onto \(\mathcal R\),
\[
  \Omega_{\mathrm{in}}^*\Omega_{\mathrm{in}}
  =
  \Omega_{\mathrm{out}}^*\Omega_{\mathrm{out}}
  =
  I_{\mathcal F_s(\Hilb_1)},\qquad
  \Omega_{\mathrm{in}}\Omega_{\mathrm{in}}^*
  =
  \Omega_{\mathrm{out}}\Omega_{\mathrm{out}}^*
  =
  P_{\mathcal R}.
\]
Consequently
\[
  S^*S
  =
  \Omega_{\mathrm{in}}^*\Omega_{\mathrm{out}}
  \Omega_{\mathrm{out}}^*\Omega_{\mathrm{in}}
  =
  \Omega_{\mathrm{in}}^*P_{\mathcal R}\Omega_{\mathrm{in}}
  =
  I_{\mathcal F_s(\Hilb_1)}
\]
using \(S^*=\Omega_{\mathrm{in}}^*\Omega_{\mathrm{out}}\), and
\[
  SS^*
  =
  \Omega_{\mathrm{out}}^*\Omega_{\mathrm{in}}
  \Omega_{\mathrm{in}}^*\Omega_{\mathrm{out}}
  =
  \Omega_{\mathrm{out}}^*P_{\mathcal R}\Omega_{\mathrm{out}}
  =
  I_{\mathcal F_s(\Hilb_1)} .
\]
Thus \(S\) is unitary.  For asymptotic vectors
\(\Phi,\Psi\in\mathcal F_s(\Hilb_1)\), the defining property of the Hilbert
adjoint gives
\[
  \langle\Phi,S\Psi\rangle
  =
  \langle\Phi,\Omega_{\mathrm{out}}^*\Omega_{\mathrm{in}}\Psi\rangle
  =
  \langle\Omega_{\mathrm{out}}\Phi,\Omega_{\mathrm{in}}\Psi\rangle_{\Hilb},
\]
which is the displayed matrix-element formula after substituting the
symmetrized tensor wave packets.

\begin{definition}[Range and completeness statements]
\label{def:hr-range-completeness-statements}
There are three logically distinct range statements.
\begin{enumerate}[label=(\alph*)]
  \item \emph{Range coincidence for the chosen asymptotic species.}  The
  equality
  \[
    \Hilb_{\mathrm{in}}=\Hilb_{\mathrm{out}}=:\Hilb_{\mathrm{sc}}
  \]
  says that the incoming and outgoing Haag--Ruelle constructions based on the
  selected isolated one-particle subspace have the same physical range.  This
  is exactly the hypothesis needed for \(S\) above to be a unitary operator on
  the corresponding asymptotic Fock space.

  \item \emph{Completeness in a specified massive scattering sector.}  Let
  \(\Hilb_{\mathrm{sec}}\subset\Hilb\) be a closed reducing subspace selected
  by superselection charge, mass gap, and a finite list of stable isolated
  particle species.  Replacing \(\mathcal F_s(\Hilb_1)\) by the appropriate
  graded Fock space over the direct sum of those one-particle spaces, sector
  completeness is the statement
  \[
    \Hilb_{\mathrm{in}}=\Hilb_{\mathrm{out}}=\Hilb_{\mathrm{sec}} .
  \]
  This is the level used by later cluster and partial-wave arguments for a
  closed massive sector: all states in that sector are accounted for by the
  listed asymptotic particles, while other superselection sectors or other
  parts of the full theory may remain outside the discussion.

  \item \emph{Full asymptotic completeness.}  After including every stable
  particle species and superselection sector in the physical theory, full
  asymptotic completeness is the statement
  \[
    \Hilb_{\mathrm{in}}=\Hilb_{\mathrm{out}}=\Hilb
  \]
  on the relevant finite-energy physical Hilbert space.
\end{enumerate}
\end{definition}

Thus ordinary asymptotic completeness is the last statement: it includes the
range coincidence in (a) and asserts that the resulting scattering sector is
all of \(\Hilb\).  The definition of \(S\) on a chosen scattering sector
requires (a).  The later closed-channel unitarity formulae require the sector
statement (b) for the sector under analysis.  Full asymptotic
completeness asserts
\[
  \Hilb=\Hilb_{\mathrm{in}}=\Hilb_{\mathrm{out}}
\]
for the whole chosen physical Hilbert space.  This is an additional dynamical
theorem or hypothesis.  For nontrivial interacting massive local Wightman
theories in four spacetime dimensions, full asymptotic completeness is
presently an open problem rather than a known consequence of the axioms.
Proofs exist only in restricted model classes, notably constructive or
integrable models in lower spacetime dimensions and related weakened-locality
settings.

\section{Relation to Local Correlation Functions}

The construction above produces the \(S\)-operator from Hilbert-space limits.
It also identifies what a later reduction formula must compute. If
\(\psi_j=P_1B_{j,t}(h_j)\vac\) and
\(\varphi_i=P_1C_{i,t}(g_i)\vac\), then an \(m\)-out, \(n\)-in matrix
element is
\[
\begin{aligned}
&\left\langle
  \Omega_{\mathrm{out}}
  (\varphi_1\odot\cdots\odot\varphi_m),\,
  \Omega_{\mathrm{in}}
  (\psi_1\odot\cdots\odot\psi_n)
  \right\rangle_{\Hilb}
\\
&\quad =
  \lim_{T\to+\infty}
  \bra{\vac}
  C_{m,T}(g_m)^*\cdots C_{1,T}(g_1)^*
  B_{1,-T}(h_1)\cdots B_{n,-T}(h_n)
  \ket{\vac}.
\end{aligned}
\]
For large \(T\), the outgoing packets are localized near future timelike
directions and the incoming packets near past timelike directions. Locality
permits these vacuum matrix elements to be reorganized into time-ordered
correlation functions up to terms that vanish in the scattering limit. The
next chapter turns this statement into LSZ reduction: one extracts the above
matrix elements from the one-particle poles of time-ordered Green functions.

\section{Scope of the Massive Construction}

The massive Haag--Ruelle construction gives a precise scattering theory for
stable particles whose one-particle mass shell is isolated and whose sector is
generated by the local creators used in the theorem.  Gauge theories require
an additional localization check.  A local gauge-invariant observable cannot
change an unscreened Gauss-law surface charge, so charged particles require
charged field algebras or noncompact gauge-invariant dressings rather than
ordinary bounded-region observable creators; the algebraic formulation is
developed in
Chapter~\ref{ch:volume-iv-haag-ruelle-mathematical-scattering}.  Massless particles, long-range
gauge forces, infraparticle sectors, and confinement then add further
asymptotic questions with their own hypotheses and limiting procedures.

The logical output of this chapter is therefore fixed. A scattering operator
has been defined before any LSZ formula and before any perturbative
interpretation of Feynman graphs as scattering amplitudes. Perturbative
scattering calculations, when available, compute matrix elements of this
operator after the reduction theorem has connected the Hilbert-space limits to
Green functions.
