\chapter{Relativistic Scalar Fields and Canonical Quantization}
\label{ch:relativistic-scalar-fields-canonical-quantization}
\ConventionsLorentzian


\section{Fields and Variational Equations}

\begin{definition}[Classical scalar Lagrangian field datum]
\label{def:classical-scalar-lagrangian-field-datum}
Let \(\mathbb M^D\) be \(D\)-dimensional Minkowski spacetime with metric
\[
  \eta_{\mu\nu}=\operatorname{diag}(-1,+1,\ldots,+1).
\]
Write
\[
  x^\mu=(x^0,\vec x),\qquad x^0=t,\qquad \vec x\in\mathbb R^{D-1}.
\]
A real classical scalar field is a real-valued function
\[
  \phi:\mathbb M^D\to\mathbb R.
\]
The field-theoretic analogue of finitely many coordinates \(q^i(t)\) is the
family of coordinates \(\phi(t,\vec x)\), labelled continuously by
\(\vec x\). A Lagrangian field theory is specified by an action functional
\[
  S[\phi]
  =
  \int_{\mathbb M^D}\dd^D x\,
  \mathcal L\bigl(\phi(x),\partial_\mu\phi(x)\bigr),
\]
where \(\mathcal L\) is a Lorentz scalar Lagrangian density in the examples
below.
\end{definition}

For a compactly supported variation \(\delta\phi\),
\[
\begin{aligned}
  \delta S
  &=
  \int\dd^Dx\,
  \left[
    \frac{\partial\mathcal L}{\partial\phi}\,\delta\phi
    +
    \frac{\partial\mathcal L}
         {\partial(\partial_\mu\phi)}
    \,\partial_\mu\delta\phi
  \right]  \\
  &=
  \int\dd^Dx\,\delta\phi
  \left[
    \frac{\partial\mathcal L}{\partial\phi}
    -
    \partial_\mu
    \frac{\partial\mathcal L}
         {\partial(\partial_\mu\phi)}
  \right].
\end{aligned}
\]
The Euler--Lagrange equation is therefore
\[
  \frac{\partial\mathcal L}{\partial\phi}
  -
  \partial_\mu
  \frac{\partial\mathcal L}{\partial(\partial_\mu\phi)}
  =
  0.
\]
The first equality is the chain rule applied pointwise to
\(\mathcal L(\phi,\partial\phi)\).  The second equality follows by integrating
\(\frac{\partial\mathcal L}{\partial(\partial_\mu\phi)}\partial_\mu\delta\phi\)
by parts.  Compact support of \(\delta\phi\), or the boundary condition class
chosen for the variational problem, removes the boundary term.  Since
\(\delta\phi\) is otherwise arbitrary, the fundamental lemma of the calculus
of variations gives the displayed Euler--Lagrange equation as a distributional
identity.

\section{The Free Massive Scalar}

\begin{definition}[Free real Klein--Gordon field]
\label{def:free-real-klein-gordon-field}
The free massive real scalar field of mass \(m\ge0\) has Lagrangian density
\[
  \mathcal L
  =
  \frac12(\partial_t\phi)^2
  -\frac12\sum_{i=1}^{D-1}(\partial_i\phi)^2
  -\frac12m^2\phi^2.
\]
Equivalently,
\[
  \mathcal L
  =
  -\frac12\partial_\mu\phi\,\partial^\mu\phi
  -\frac12m^2\phi^2,
  \qquad
  \partial^\mu=\eta^{\mu\nu}\partial_\nu.
\]
The Euler--Lagrange equation is the Klein--Gordon equation
\[
  (\partial^\mu\partial_\mu-m^2)\phi=0.
\]
\end{definition}

\begin{proposition}[Mass-shell support of free solutions]
\label{prop:kg-mass-shell-support}
Use the Fourier convention
\[
  \phi(x)
  =
  \int\frac{\dd^Dk}{(2\pi)^D}\,
  \widetilde\phi(k)\,\ee^{\ii k\cdot x},
  \qquad
  k\cdot x=\eta_{\mu\nu}k^\mu x^\nu
  =
  -k^0x^0+\vec k\cdot\vec x.
\]
The equation of motion implies
\[
  (k^2+m^2)\widetilde\phi(k)=0,
  \qquad
  k^2=\eta_{\mu\nu}k^\mu k^\nu.
\]
Thus \(\widetilde\phi(k)\) is supported on the mass shell
\[
  k^2+m^2=0,
  \qquad
  k^0=\pm\omega_{\vec k},
  \qquad
  \omega_{\vec k}=\sqrt{\vec k^{\,2}+m^2}.
\]
\end{proposition}

\begin{proof}
With \(k\cdot x=-k^0x^0+\vec k\cdot\vec x\), one has
\[
  \partial^\mu\partial_\mu \ee^{\ii k\cdot x}
  =
  -\bigl(k^2\bigr)\ee^{\ii k\cdot x}.
\]
The Klein--Gordon equation therefore becomes
\(-(k^2+m^2)\widetilde\phi(k)=0\) in Fourier space.  In distribution
language this says that \(\widetilde\phi\) is supported where
\(k^2+m^2=0\).  Since \(k^2=-(k^0)^2+\vec k^{\,2}\), this hypersurface is
\((k^0)^2=\vec k^{\,2}+m^2\), hence the two branches
\(k^0=\pm\omega_{\vec k}\).
\end{proof}

Figure~\ref{fig:scalar-mass-shell-branches} displays the two branches of this
support condition and fixes the positive-frequency branch used in the canonical
expansion below.

\begin{figure}[H]
\centering
\begin{tikzpicture}[>=Stealth,scale=0.95]
  \draw[->] (-3.15,0) -- (3.35,0) node[right] {$|\vec k|$};
  \draw[->] (0,-2.55) -- (0,2.75) node[above] {$k^0$};
  \draw[blue!70!black,line width=1pt,domain=-2.65:2.65,samples=100]
    plot (\x,{sqrt(0.65+\x*\x)});
  \draw[blue!70!black,line width=1pt,domain=-2.65:2.65,samples=100]
    plot (\x,{-sqrt(0.65+\x*\x)});
  \node[blue!70!black,above right] at (1.85,2.55) {$k^0=+\omega_{\vec k}$};
  \node[blue!70!black,below right] at (1.85,-2.55) {$k^0=-\omega_{\vec k}$};
  \node at (-1.25,0.35) {$k^2+m^2=0$};
\end{tikzpicture}
\caption{Mass shell for a real scalar solution.  The Fourier transform
\(\widetilde\phi(k)\) is supported on the two branches
\(k^0=\pm\omega_{\vec k}\) of \(k^2+m^2=0\).}
\label{fig:scalar-mass-shell-branches}
\end{figure}

\begin{proposition}[Cauchy data determine the free solution]
\label{prop:kg-cauchy-data-determine-free-solution}
A real solution can be written as
\[
  \phi(t,\vec x)
  =
  \int\frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}}
  \left[
    A(\vec k)\,\ee^{\ii\vec k\cdot\vec x-\ii\omega_{\vec k}t}
    +
    A(\vec k)^*\,\ee^{-\ii\vec k\cdot\vec x+\ii\omega_{\vec k}t}
  \right],
\]
with a complex coefficient function \(A(\vec k)\) subject to whatever
regularity conditions are imposed on the classical solution. If \(m=0\), the
spatial zero mode \(\vec k=0\) is an infrared mode and is treated separately
by the chosen boundary condition, volume regulator, or test-function class;
the displayed division by \(\omega_{\vec k}\) applies on the modes with
\(\omega_{\vec k}>0\).  The Cauchy data
\[
  \phi(0,\vec x),\qquad \partial_t\phi(0,\vec x)
\]
determine \(A(\vec k)\) and \(A(\vec k)^*\), hence determine the free
solution for all \(t\). Explicitly, write
\[
  \phi(0,\vec x)
  =
  \int\frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}}\,
  \widetilde\phi_0(\vec k)\ee^{\ii\vec k\cdot\vec x},
  \qquad
  \partial_t\phi(0,\vec x)
  =
  \int\frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}}\,
  \widetilde\pi_0(\vec k)\ee^{\ii\vec k\cdot\vec x}.
\]
Then
\[
  \widetilde\phi_0(\vec k)
  =
  A(\vec k)+A(-\vec k)^*,
  \qquad
  \widetilde\pi_0(\vec k)
  =
  -\ii\omega_{\vec k}\bigl(A(\vec k)-A(-\vec k)^*\bigr),
\]
so that
\[
  A(\vec k)
  =
  \frac12\left(
    \widetilde\phi_0(\vec k)
    +
    \frac{\ii}{\omega_{\vec k}}\widetilde\pi_0(\vec k)
  \right).
\]
\end{proposition}

\begin{proof}
Evaluate the positive- and negative-frequency expansion at \(t=0\) and also
differentiate it once in time before setting \(t=0\).  Reality of \(\phi\)
identifies the coefficient of \(\ee^{-\ii\vec k\cdot\vec x}\) with the
complex conjugate coefficient at momentum \(\vec k\), equivalently
\(A(-\vec k)^*\) in the \(\ee^{\ii\vec k\cdot\vec x}\) coefficient.  This
gives the two displayed linear equations for
\(A(\vec k)\) and \(A(-\vec k)^*\).  Adding
\(\widetilde\phi_0(\vec k)\) to \(\ii\widetilde\pi_0(\vec k)/\omega_{\vec k}\)
solves the system and gives the stated formula.  The Klein--Gordon equation is
second order in time, so these two Cauchy data determine the solution on every
time slice.
\end{proof}

There are two constructions to keep distinct. Canonical quantization starts
from the equal-time phase space \((\phi,\Pi)\) and a Hamiltonian. The
Lagrangian path-integral construction is defined by first choosing a regulator
that replaces the field by finitely many coordinates or modes:
\[
  \int_{\mathcal C_R}\dd\mu_R(\phi_R)\,
  \exp\left(\frac{\ii}{\hbar}S_R[\phi_R]\right).
\]
Here \(R\) labels the chosen cutoff, \(\mathcal C_R\) is the regulated
configuration space, \(\dd\mu_R\) is the corresponding finite-dimensional or
lattice measure, and \(S_R\) is the regulated action.  With the Lorentzian
phase \(\exp(\ii S_R/\hbar)\), this regulated object is an oscillatory
integral or oscillatory distribution.  For quadratic actions it is a Fresnel
integral with a determinant magnitude and a phase fixed by the signature of the
quadratic form; for general actions, stationary phase and distributional
regularization are part of the definition.  The continuum symbol
\(\int[D\phi]\exp(\ii S[\phi]/\hbar)\) denotes a specified limit or asymptotic
expansion of these regulated oscillatory objects. This chapter develops the
canonical construction; the scalar field path integral is introduced later.

\section{Canonical Hamiltonian Formalism}

\begin{construction}[Canonical phase-space variables and Hamiltonian density]
Let
\[
  L(t)=\int\dd^{D-1}\vec x\,\mathcal L.
\]
The canonical momentum density conjugate to \(\phi(t,\vec x)\) is
\[
  \Pi(t,\vec x)
  =
  \frac{\partial\mathcal L}
       {\partial(\partial_t\phi(t,\vec x))}.
\]
For the free scalar,
\[
  \Pi(t,\vec x)=\partial_t\phi(t,\vec x).
\]
This agrees with the covariant form of the Lagrangian.  Treating the
components \(\partial_\mu\phi\) as independent variables in
\(-\frac12\eta^{\alpha\beta}\partial_\alpha\phi\,\partial_\beta\phi\), one
finds
\[
  \frac{\partial\mathcal L}{\partial(\partial_0\phi)}
  =
  -\eta^{0\nu}\partial_\nu\phi
  =
  -\partial^0\phi
  =
  \partial_0\phi
  =
  \partial_t\phi ,
\]
because the mostly-plus metric gives
\(\partial^0=\eta^{00}\partial_0=-\partial_0\).  The two minus signs,
one from the covariant kinetic term and one from raising the time index,
therefore cancel.
The Hamiltonian density is
\[
  \mathcal H
  =
  \Pi\,\partial_t\phi-\mathcal L,
\]
and therefore
\[
  \mathcal H
  =
  \frac12\Pi^2
  +
  \frac12\sum_{i=1}^{D-1}(\partial_i\phi)^2
  +
  \frac12m^2\phi^2
\]
for the free scalar field. The Hamiltonian is
\[
  H=\int\dd^{D-1}\vec x\,\mathcal H.
\]
\end{construction}

This is a classical Hamiltonian density.  After quantization
\(\widehat\phi(t,\vec x)\) and \(\widehat\Pi(t,\vec x)\) are
operator-valued distributions, so the pointwise products in
\(\frac12(\widehat\Pi^2+(\nabla\widehat\phi)^2+m^2\widehat\phi^2)\) are first
understood with a spatial regulator or after smearing.  The free Hamiltonian
used below is the regulator limit of the normal-ordered quadratic Hamiltonian,
equivalently the operator obtained by subtracting the vacuum scalar term from
the regulated quadratic expression.  We write this operator as
\(\widehat H_0^{\mathrm{ren}}=:\widehat H_0:\) in the flat-space free theory.

\begin{definition}[Equal-time scalar Poisson bracket]
\label{def:equal-time-scalar-poisson-bracket}
For functionals \(F[\phi,\Pi]\) and \(G[\phi,\Pi]\), use the equal-time
Poisson bracket
\[
  \{F,G\}
  =
  \int\dd^{D-1}\vec x
  \left[
    \frac{\delta F}{\delta\phi(\vec x)}
    \frac{\delta G}{\delta\Pi(\vec x)}
    -
    \frac{\delta F}{\delta\Pi(\vec x)}
    \frac{\delta G}{\delta\phi(\vec x)}
  \right].
\]
Then
\[
  \{\phi(\vec x),\Pi(\vec y)\}
  =
  \delta^{(D-1)}(\vec x-\vec y),
\]
with the \(\phi\)-\(\phi\) and \(\Pi\)-\(\Pi\) brackets equal to zero. Hamilton's
equation is
\[
  \partial_tF=\{F,H\}.
\]
\end{definition}

\paragraph{Hamilton equations for the scalar field.}
Using the functional derivatives of \(H\),
\[
  \frac{\delta H}{\delta\Pi(\vec x)}=\Pi(\vec x),
  \qquad
  \frac{\delta H}{\delta\phi(\vec x)}
  =
  -\sum_i\partial_i^2\phi(\vec x)+m^2\phi(\vec x),
\]
where the spatial derivative term has been integrated by parts.  Hence
\[
  \partial_t\phi=\{\phi,H\}=\Pi,
  \qquad
  \partial_t\Pi=\{\Pi,H\}
  =
  \sum_i\partial_i^2\phi-m^2\phi.
\]
Eliminating \(\Pi\) gives
\(\partial_t^2\phi-\nabla^2\phi+m^2\phi=0\), equivalently
\((\partial^\mu\partial_\mu-m^2)\phi=0\) with the mostly-plus convention.

\section{Equal-Time Quantization}

\begin{definition}[Equal-time canonical commutation relations]
\label{def:equal-time-scalar-ccr}
Canonical quantization promotes the classical fields to operator-valued
distributions
\[
  \widehat\phi(t,\vec x),\qquad \widehat\Pi(t,\vec x)
\]
on a common dense domain. Their equal-time commutation relations are
\[
  [\widehat\phi(t,\vec x),\widehat\Pi(t,\vec y)]
  =
  \ii\hbar\,\delta^{(D-1)}(\vec x-\vec y),
\]
and
\[
  [\widehat\phi(t,\vec x),\widehat\phi(t,\vec y)]
  =
  0
  =
  [\widehat\Pi(t,\vec x),\widehat\Pi(t,\vec y)].
\]
All pointwise products are distributional notation. Smeared fields, for
test functions \(f,g\in C_c^\infty(\mathbb R^{D-1})\), are
\[
  \widehat\phi_t(f)=\int\dd^{D-1}\vec x\,f(\vec x)\widehat\phi(t,\vec x),
  \qquad
  \widehat\Pi_t(g)=\int\dd^{D-1}\vec x\,g(\vec x)\widehat\Pi(t,\vec x).
\]
The smeared commutator is
\[
  [\widehat\phi_t(f),\widehat\Pi_t(g)]
  =
  \ii\hbar
  \int\dd^{D-1}\vec x\,f(\vec x)g(\vec x).
\]
\end{definition}

For the rest of this chapter set \(\hbar=1\). With \(\hbar\) retained, the
oscillator algebra below reads
\[
  [b_{\vec k},b^\dagger_{\vec q}]
  =
  \hbar(2\pi)^{D-1}\delta^{(D-1)}(\vec k-\vec q).
\]
Equivalently, the \(\hbar=1\) variables are obtained from the \(\hbar\)-full
variables by \(b_{\vec k}^{(\hbar=1)}=\hbar^{-1/2}b_{\vec k}^{(\hbar)}\).

\section{Oscillator Variables and the Hamiltonian}

\begin{proposition}[Equal-time oscillator variables and the CCR]
\label{prop:equal-time-scalar-oscillator-variables-ccr}
Let \(\nu_{\vec k}>0\) be a positive even function of \(\vec k\). Define
operator-valued distributions \(b_{\vec k}\) and \(b^\dagger_{\vec k}\) by the
equal-time expansion
\[
  \widehat\phi(0,\vec x)
  =
  \int\frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}}\,
  \frac{1}{\sqrt{2\nu_{\vec k}}}
  \left(
    b_{\vec k}\ee^{\ii\vec k\cdot\vec x}
    +
    b^\dagger_{\vec k}\ee^{-\ii\vec k\cdot\vec x}
  \right),
\]
\[
  \widehat\Pi(0,\vec x)
  =
  \int\frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}}\,
  (-\ii)\sqrt{\frac{\nu_{\vec k}}{2}}
  \left(
    b_{\vec k}\ee^{\ii\vec k\cdot\vec x}
    -
    b^\dagger_{\vec k}\ee^{-\ii\vec k\cdot\vec x}
  \right).
\]
The equal-time commutation relations are equivalent to
\[
  [b_{\vec k},b_{\vec q}]=0
  =
  [b^\dagger_{\vec k},b^\dagger_{\vec q}],
  \qquad
  [b_{\vec k},b^\dagger_{\vec q}]
  =
  (2\pi)^{D-1}\delta^{(D-1)}(\vec k-\vec q).
\]
\end{proposition}

\begin{proof}
Substitute the mode expansions into the smeared equal-time commutator and use
\[
  \int\frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}}\,
  \ee^{\ii\vec k\cdot(\vec x-\vec y)}
  =
  \delta^{(D-1)}(\vec x-\vec y).
\]
The coefficient
\((2\nu_{\vec k})^{-1/2}(-\ii)(\nu_{\vec k}/2)^{1/2}\), together with the
term obtained by commuting \(b^\dagger\) past \(b\), gives exactly
\(\ii\delta^{(D-1)}(\vec x-\vec y)\) when the oscillator commutator has the
displayed normalization.  Conversely, Fourier transforming the canonical
commutator against plane waves recovers the oscillator commutator.  The
\(\phi\)-\(\phi\) and \(\Pi\)-\(\Pi\) commutators vanish because the
\(b\)-\(b\) and \(b^\dagger\)-\(b^\dagger\) commutators vanish and the
remaining terms cancel by \(\vec k\mapsto-\vec k\).
\end{proof}

\begin{proposition}[Bogoliubov reparametrizations preserve the CCR]
\label{prop:scalar-bogoliubov-reparametrizations-preserve-ccr}
This algebra holds for every positive choice of \(\nu_{\vec k}\). The
notation \(b_{\vec k}\), \(b^\dagger_{\vec k}\) should therefore not yet be
read as a particle interpretation. If \(\lambda_{\vec k}>0\) is another even
function, define \(c_{\vec k}\), \(c^\dagger_{\vec k}\) by
\[
  c_{\vec k}+c^\dagger_{-\vec k}
  =
  \lambda_{\vec k}\bigl(b_{\vec k}+b^\dagger_{-\vec k}\bigr),
  \qquad
  c_{\vec k}-c^\dagger_{-\vec k}
  =
  \lambda_{\vec k}^{-1}\bigl(b_{\vec k}-b^\dagger_{-\vec k}\bigr).
\]
Equivalently,
\[
  c_{\vec k}
  =
  \frac{\lambda_{\vec k}+\lambda_{\vec k}^{-1}}{2}\,b_{\vec k}
  +
  \frac{\lambda_{\vec k}-\lambda_{\vec k}^{-1}}{2}\,
  b^\dagger_{-\vec k}.
\]
The \(c\)'s obey the same canonical oscillator algebra. If a vector is
annihilated by every \(b_{\vec k}\), it is not annihilated by every
\(c_{\vec k}\) unless \(\lambda_{\vec k}=1\). The Hamiltonian is the datum
that selects the oscillator normalization.
\end{proposition}

\begin{proof}
Write \(u_{\vec k}=(\lambda_{\vec k}+\lambda_{\vec k}^{-1})/2\) and
\(v_{\vec k}=(\lambda_{\vec k}-\lambda_{\vec k}^{-1})/2\).  Evenness of
\(\lambda\) gives \(u_{-\vec k}=u_{\vec k}\) and \(v_{-\vec k}=v_{\vec k}\).
The only nonzero commutators in
\([c_{\vec k},c^\dagger_{\vec q}]\) are
\[
  u_{\vec k}u_{\vec q}[b_{\vec k},b^\dagger_{\vec q}]
  -
  v_{\vec k}v_{\vec q}[b_{-\vec q},b^\dagger_{-\vec k}],
\]
and the delta functions set \(\vec k=\vec q\).  Since
\(u_{\vec k}^2-v_{\vec k}^2=1\), the original oscillator normalization is
preserved.  The \(c\)-\(c\) and \(c^\dagger\)-\(c^\dagger\) commutators vanish
by the same cancellation.  Finally, if \(b_{\vec k}\vac_b=0\), then
\[
  c_{\vec k}\vac_b
  =
  v_{\vec k}b^\dagger_{-\vec k}\vac_b,
\]
which vanishes for all \(\vec k\) only when \(v_{\vec k}=0\), equivalently
\(\lambda_{\vec k}=1\).
\end{proof}

\begin{proposition}[Free Hamiltonian selects \(\nu_{\vec k}=\omega_{\vec k}\)]
\label{prop:free-hamiltonian-selects-omega-normalization}
Let
\[
  \omega_{\vec k}=\sqrt{\vec k^{\,2}+m^2}.
\]
Substituting the mode expansion into the free Hamiltonian gives
\[
\begin{aligned}
  \widehat H_0
  =
  \int\frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}}
  \Bigg[
  &\frac{\nu_{\vec k}^2+\omega_{\vec k}^2}{2\nu_{\vec k}}
  \left(
    b^\dagger_{\vec k}b_{\vec k}
    +\frac12(2\pi)^{D-1}\delta^{(D-1)}(0)
  \right)\\
  &+
  \frac{\omega_{\vec k}^2-\nu_{\vec k}^2}{4\nu_{\vec k}}
  \left(
    b_{\vec k}b_{-\vec k}
    +
    b^\dagger_{\vec k}b^\dagger_{-\vec k}
  \right)
  \Bigg].
\end{aligned}
\]
The off-diagonal terms vanish precisely for
\[
  \nu_{\vec k}=\omega_{\vec k}.
\]
\end{proposition}

\begin{proof}
Insert the equal-time oscillator expansion into
\[
  H_0=\frac12\int\dd^{D-1}\vec x\,
  \bigl(\Pi^2+(\nabla\phi)^2+m^2\phi^2\bigr).
\]
The spatial integral pairs momenta \(\vec k\) and \(-\vec k\).  The
\(\Pi^2\) term contributes \(\nu_{\vec k}/2\) to
\(b^\dagger_{\vec k}b_{\vec k}\) and \(-\nu_{\vec k}/4\) to the
\(b_{\vec k}b_{-\vec k}+b^\dagger_{\vec k}b^\dagger_{-\vec k}\) coefficient,
while the gradient and mass terms together contribute
\(\omega_{\vec k}^2/(2\nu_{\vec k})\) to
\(b^\dagger_{\vec k}b_{\vec k}\) and
\(\omega_{\vec k}^2/(4\nu_{\vec k})\) to the off-diagonal coefficient.  Adding
the terms gives the displayed Hamiltonian.  Since \(\nu_{\vec k}>0\), the
off-diagonal coefficient vanishes if and only if
\(\nu_{\vec k}^2=\omega_{\vec k}^2\), hence
\(\nu_{\vec k}=\omega_{\vec k}\).
\end{proof}

\begin{definition}[Free scalar Fock representation]
\label{def:free-scalar-fock-representation}
With this choice,
\[
  \widehat H_0
  =
  \int\frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}}\,
  \omega_{\vec k}
  \left(
    b^\dagger_{\vec k}b_{\vec k}
    +\frac12(2\pi)^{D-1}\delta^{(D-1)}(0)
  \right).
\]
At finite spatial volume and finite momentum cutoff, the second term is a
finite scalar multiple of the identity. A constant multiple of the identity
does not change Heisenberg evolution, but the quantum Hamiltonian must be a
well-defined self-adjoint operator after the regulator is removed. We
therefore choose the additive energy normalization so that the invariant
vacuum has zero energy. This is a convention of the flat-spacetime framework
with gravity decoupled.  If the metric is dynamical, the vacuum expectation of
the stress tensor contributes to the cosmological constant, and its finite
renormalized value is additional physical data rather than a consequence of
normal ordering. With this normalization the free Hamiltonian is
\[
  \widehat H_0^{\mathrm{ren}}
  =
  \int\frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}}\,
  \omega_{\vec k} b^\dagger_{\vec k}b_{\vec k}.
\]
The free vacuum \(\vac\) is defined by
\[
  b_{\vec k}\vac=0
  \qquad\text{for all }\vec k,
\]
and satisfies \(\widehat H_0^{\mathrm{ren}}\vac=0\).
\end{definition}

\begin{proposition}[Heisenberg field and invariant mass-shell normalization]
\label{prop:heisenberg-field-invariant-mass-shell-normalization}
Time evolution generated by \(\widehat H_0^{\mathrm{ren}}\) gives
\[
  \widehat\phi(x)
  =
  \int\frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}}\,
  \frac{1}{\sqrt{2\omega_{\vec k}}}
  \left(
    b_{\vec k}\ee^{\ii\vec k\cdot\vec x-\ii\omega_{\vec k}t}
    +
    b^\dagger_{\vec k}\ee^{-\ii\vec k\cdot\vec x+\ii\omega_{\vec k}t}
  \right).
\]

The invariant mass-shell normalization used earlier is obtained by defining
\[
  a(\vec k)=\sqrt{2\omega_{\vec k}}\,b_{\vec k},
  \qquad
  a^\dagger(\vec k)=\sqrt{2\omega_{\vec k}}\,b^\dagger_{\vec k}.
\]
Then
\[
  [a(\vec k),a^\dagger(\vec q)]
  =
  (2\pi)^{D-1}2\omega_{\vec k}\,
  \delta^{(D-1)}(\vec k-\vec q),
\]
and
\[
  \widehat\phi(x)
  =
  \int
  \frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}2\omega_{\vec k}}\,
  \left(
    a(\vec k)\ee^{\ii k\cdot x}
    +
    a^\dagger(\vec k)\ee^{-\ii k\cdot x}
  \right),
  \qquad
  k^0=\omega_{\vec k}.
\]
This is the convention used for the free scalar field in the earlier
Fock-space construction.
\end{proposition}

\begin{proof}
The commutator with the normal-ordered Hamiltonian gives
\[
  [\widehat H_0^{\rm ren},b_{\vec k}]=-\omega_{\vec k}b_{\vec k},
  \qquad
  [\widehat H_0^{\rm ren},b^\dagger_{\vec k}]
  =
  \omega_{\vec k}b^\dagger_{\vec k},
\]
so Heisenberg evolution multiplies \(b_{\vec k}\) by
\(\ee^{-\ii\omega_{\vec k}t}\) and \(b^\dagger_{\vec k}\) by
\(\ee^{+\ii\omega_{\vec k}t}\).  This gives the displayed field.  The
rescaling \(a(\vec k)=\sqrt{2\omega_{\vec k}}\,b_{\vec k}\) transforms the
Lebesgue momentum measure into the invariant one-particle mass-shell measure
\(\dd^{D-1}\vec k/((2\pi)^{D-1}2\omega_{\vec k})\) and gives the stated
commutator for \(a,a^\dagger\).
\end{proof}

\section{Covariance and Microcausality}

\begin{definition}[Free scalar field on the finite-particle domain]
\label{def:free-scalar-field-finite-particle-domain}
Let \(\mathcal D_0\) be the finite-particle domain generated by applying
finitely many smeared creation operators to \(\vac\). On \(\mathcal D_0\),
\(\widehat\phi(x)\) is an operator-valued distribution. For
\(f\in C_c^\infty(\mathbb M^D)\),
\[
  \widehat\phi(f)=\int\dd^D x\,f(x)\widehat\phi(x)
\]
is the corresponding smeared operator on \(\mathcal D_0\).
\end{definition}

\begin{proposition}[Covariance and microcausality of the free scalar field]
\label{prop:free-scalar-covariance-microcausality}
The free Fock representation carries the unitary Poincare representation
induced from the one-particle mass shell. With the above normalization, the
field transforms as a scalar:
\[
  U(a,\Lambda)^{-1}\widehat\phi(x)U(a,\Lambda)
  =
  \widehat\phi(\Lambda x+a).
\]

The commutator is
\[
  [\widehat\phi(x),\widehat\phi(y)]
  =
  \ii\,\Delta(x-y),
\]
where the Pauli--Jordan distribution is
\[
  \Delta(x)
  =
  \frac{1}{\ii}
  \int
  \frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}2\omega_{\vec k}}
  \left(
    \ee^{\ii k\cdot x}
    -
    \ee^{-\ii k\cdot x}
  \right),
  \qquad k^0=\omega_{\vec k}.
\]
This distribution is Lorentz invariant. If \(x\) is spacelike, a Lorentz
transformation sends it to \((0,\vec r)\). At equal time,
\[
  \Delta(0,\vec r)
  =
  \frac{1}{\ii}
  \int
  \frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}2\omega_{\vec k}}
  \left(
    \ee^{\ii\vec k\cdot\vec r}
    -
    \ee^{-\ii\vec k\cdot\vec r}
  \right)
  =
  0,
\]
because the two terms are exchanged by \(\vec k\mapsto-\vec k\). Therefore
\[
  [\widehat\phi(x),\widehat\phi(y)]=0
  \qquad
  \text{when }(x-y)^2>0.
\]
Equivalently,
\[
  [\widehat\phi(f),\widehat\phi(g)]=0
\]
when \(\operatorname{supp}f\) and \(\operatorname{supp}g\) are spacelike
separated. This verifies microcausality for the canonically quantized free
scalar field.
\end{proposition}

\begin{proof}
The one-particle Hilbert space is
\[
  L^2\!\left(\{k^2+m^2=0,\ k^0>0\},
  \frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}2\omega_{\vec k}}\right),
\]
and the positive-energy mass-shell measure is Lorentz invariant.  The induced
unitary representation acts by translating phases and by pulling back the
mass-shell momentum under Lorentz transformations; second quantization gives
the Fock-space representation.  Substituting this action into the mass-shell
expansion of \(\widehat\phi(x)\) gives the displayed scalar covariance law.

The commutator computation uses only
\([a(\vec k),a^\dagger(\vec q)]=(2\pi)^{D-1}2\omega_{\vec k}
\delta^{(D-1)}(\vec k-\vec q)\), and yields
\([\widehat\phi(x),\widehat\phi(y)]=\ii\Delta(x-y)\).  The distribution
\(\Delta\) is Lorentz invariant because the mass-shell measure and the
condition \(k^0>0\) are invariant under proper orthochronous Lorentz
transformations, and the difference between the two exponentials gives the
antisymmetric solution of the Klein--Gordon equation.  If \(x-y\) is
spacelike, a Lorentz transformation sends it to an equal-time vector
\((0,\vec r)\).  The equal-time integral vanishes by the change of variables
\(\vec k\mapsto-\vec k\).  Thus \(\Delta(x-y)=0\) for spacelike separation.
After smearing, the commutator is the pairing of \(\Delta(x-y)\) with
\(f(x)g(y)\), so it vanishes whenever the two supports are spacelike
separated.
\end{proof}

\begin{remark}
For the scalar field, locality is carried by the commutator.  The Pauli--Jordan
distribution is the difference of the positive-frequency two-point distribution
and its reversed argument, and this difference is supported on the causal cone.
The anticommutator has the symmetric vacuum distribution
\[
  \Delta_+(x-y)+\Delta_+(y-x),
\]
whose equal-time spacelike kernel is proportional to
\[
  \int
  \frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}2\omega_{\vec k}}\,
  \cos(\vec k\cdot\vec r).
\]
This distribution is not supported in the causal cone and generally has
nonzero spacelike matrix elements.  The graded sign in the locality bracket is
therefore part of the structure fixed by spin and positivity, rather than an
optional convention.  The corresponding spin-statistics theorem-level
hypotheses are stated in Chapter~\ref{ch:spinor-fields-fermionic-path-integrals}.
\end{remark}

The next step in the construction is to identify the local currents and
charges attached to spacetime symmetries. That is the role of Noether's
theorem in the following chapter.
