\chapter{Quantum Electrodynamics and External States}
\label{ch:qed-external-states}
\ConventionsLorentzian


\section{Charged Spinors and Abelian Gauge Representatives}

\begin{definition}[QED representative datum]
\label{def:qed-representative-datum}
A local representative datum for four-dimensional QED consists of a real
Abelian vector-potential representative \(A_\mu(x)\), a Dirac spinor
representative \(\psi_\alpha(x)\), its mostly-plus Dirac adjoint
\(\bar\psi^\alpha(x)=\psi^\dagger\beta\), and a real charge coordinate
\(g\in\mathbb R\).  For the electron convention used below,
\[
  g=-e,\qquad e>0 .
\]
The representative fields are therefore
\[
  A_\mu(x),\qquad \psi_\alpha(x),\qquad \bar\psi^\alpha(x),
\]
with \(\bar\psi\) defined as in
Chapter~\ref{ch:spinor-fields-fermionic-path-integrals}.  A real function
\(\xi(x)\) changes representatives by
\[
  A_\mu\mapsto A_\mu+\partial_\mu\xi,\qquad
  \psi\mapsto \ee^{\ii g\xi}\psi,\qquad
  \bar\psi\mapsto \bar\psi\,\ee^{-\ii g\xi}.
\]
The Abelian covariant derivative is
\[
  D_\mu\psi
  =
  \partial_\mu\psi-\ii gA_\mu\psi .
\]
\end{definition}

\begin{proposition}[Covariance of the Abelian Dirac coupling]
\label{prop:qed-abelian-dirac-coupling-covariant}
For the representative change of
Definition~\ref{def:qed-representative-datum}, the covariant derivative
transforms as
\[
  D_\mu(\ee^{\ii g\xi}\psi)=\ee^{\ii g\xi}D_\mu\psi .
\]
Consequently the local action
\[
  S_{\mathrm{QED}}[A,\bar\psi,\psi]
  =
  \int\dd^4x\,
  \left[
    -\frac14F_{\mu\nu}F^{\mu\nu}
    -\bar\psi(\gamma^\mu D_\mu+m)\psi
  \right]
\]
is invariant under compactly supported changes of Abelian representative,
where
\[
  F_{\mu\nu}=\partial_\mu A_\nu-\partial_\nu A_\mu,\qquad
  \not a=a_\mu\gamma^\mu .
\]
In the monograph's mostly-plus gamma convention the current represented by
the coupling to \(A_\mu\) is
\[
  j^\mu=\ii g\,\bar\psi\gamma^\mu\psi .
\]
\end{proposition}

\begin{proof}
It transforms with the same charge as \(\psi\):
\[
\begin{aligned}
  D_\mu(\ee^{\ii g\xi}\psi)
  &=
  \partial_\mu(\ee^{\ii g\xi}\psi)
  -\ii g(A_\mu+\partial_\mu\xi)\ee^{\ii g\xi}\psi  \\
  &=
  \ee^{\ii g\xi}D_\mu\psi .
\end{aligned}
\]
The curvature \(F_{\mu\nu}\) is unchanged because
\(\partial_\mu\partial_\nu\xi=\partial_\nu\partial_\mu\xi\).  The Dirac term
is invariant because the phase of \(D_\mu\psi\) cancels the opposite phase of
\(\bar\psi\).  Expanding the covariant derivative gives
\[
  \mathcal L_{\mathrm{QED}}
  =
  -\frac14F_{\mu\nu}F^{\mu\nu}
  -\bar\psi(\not\partial+m)\psi
  +\ii gA_\mu\bar\psi\gamma^\mu\psi .
\]
The last term is the interaction. In these gamma-matrix conventions the
electric current represented by the coupling to \(A_\mu\) is
\[
  j^\mu=\ii g\,\bar\psi\gamma^\mu\psi .
\]
The leading factor of \(\ii\) is fixed by the mostly-plus Dirac adjoint
\(\bar\psi=\psi^\dagger\beta\), \(\beta=\ii\gamma^0\), together with
\(D_\mu=\partial_\mu-\ii gA_\mu\).
It is invariant under the local Abelian representative change because the
endpoint phases of \(\bar\psi\) and \(\psi\) cancel.
\end{proof}

\section{Local Observables and Charged Insertions}

\begin{definition}[Local Abelian charge of a representative monomial]
\label{def:qed-local-abelian-charge}
A local observable is represented by a local expression in the fields that is
invariant under all compactly supported changes of gauge representative.
Examples are
\[
  F_{\mu\nu},\qquad
  \bar\psi\psi,\qquad
  \ii g\,\bar\psi\gamma^\mu\psi .
\]
A local monomial containing \(N_\psi\) factors of \(\psi\) and
\(N_{\bar\psi}\) factors of \(\bar\psi\) has local charge
\[
  g(N_\psi-N_{\bar\psi}) .
\]
\end{definition}

\begin{remark}[Neutrality criterion for pointlike local monomials]
\label{prop:qed-pointlike-local-neutrality}
Gauge invariance of a pointlike local monomial therefore requires the net
local charge in Definition~\ref{def:qed-local-abelian-charge} to vanish.
Under a compactly supported representative change the monomial is multiplied
by
\[
  \exp\!\left(\ii g(N_\psi-N_{\bar\psi})\xi(x)\right)
\]
at the point of insertion.  Since \(\xi(x)\) can be chosen arbitrarily while
remaining compactly supported, invariance for all representative changes
forces \(g(N_\psi-N_{\bar\psi})=0\).  For nonzero charge coordinate \(g\) this
is precisely \(N_\psi=N_{\bar\psi}\).
\end{remark}

\begin{definition}[Wilson-line dressed charged insertion]
\label{def:qed-wilson-line-dressed-charged-insertion}
Charged insertions are represented by adding dressing data. Let \(C_x\) be an
oriented curve beginning at \(x\) and ending at a prescribed asymptotic
endpoint. For gauge parameters that vanish at that endpoint, set
\[
  \Psi_{C_x}(x)
  =
  \exp\left(\ii g\int_{C_x} A_\mu(y)\,\dd y^\mu\right)\psi(x).
\]
Similarly, for a path \(C_{y\to x}\) from \(y\) to \(x\), set
\[
  \mathcal O_{\alpha}{}^{\beta}(x,y;C_{y\to x})
  =
  \psi_\alpha(x)
  \exp\left(
    -\ii g\int_{C_{y\to x}} A_\mu(z)\,\dd z^\mu
  \right)
  \bar\psi^\beta(y).
\]
The curve is part of the insertion datum; these are not pointlike local
observables.
\end{definition}

\begin{remark}[Endpoint cancellation for Abelian Wilson dressings]
\label{prop:qed-wilson-dressing-endpoint-cancellation}
The dressed operators in
Definition~\ref{def:qed-wilson-line-dressed-charged-insertion} are invariant
under compactly supported Abelian representative changes compatible with the
stated endpoint convention.
For the half-infinite dressing,
\[
  \int_{C_x}\partial_\mu\xi(y)\,\dd y^\mu
  =
  \xi(\infty)-\xi(x)
  =
  -\xi(x),
\]
so the phase acquired by the line integral cancels the phase of \(\psi(x)\).
For the finite bilocal dressing the Wilson-line phase changes by
\[
  \exp\{-\ii g(\xi(x)-\xi(y))\}.
\]
This cancels the endpoint phases of \(\psi_\alpha(x)\) and
\(\bar\psi^\beta(y)\).
\end{remark}

The line operator itself may require short-distance
renormalization when it is defined as an operator-valued distribution. That
renormalization question is separate from the finite gauge-covariance
calculation displayed here.

\begin{remark}
Gauge-fixed perturbation theory uses \(A_\mu,\psi,\bar\psi\) Green functions
as representative correlators. After gauge fixing these correlators have
propagators and vertices. Gauge-invariant local observables and physical
external photon states are read from \(F_{\mu\nu}\), helicity representatives,
and LSZ pole data.
\end{remark}

\section{Gauge-Fixed Generating Functional}

For perturbation theory choose the covariant gauge parameter
\(\xi_{\mathrm g}\) and add
\[
  \Delta\mathcal L
  =
  -\frac{1}{2\xi_{\mathrm g}}(\partial_\mu A^\mu)^2 .
\]
The Abelian Faddeev--Popov determinant for the condition
\(\partial_\mu A^\mu=\varphi\) is \(\det\Box\), independent of \(A_\mu\);
it is absorbed into the normalization of the generating functional.

\paragraph{Abelian Faddeev--Popov specialization.}
Let
\[
  F_\varphi[A](x)=\partial_\mu A^\mu(x)-\varphi(x)
\]
be the covariant gauge condition, with \(\varphi\) an auxiliary scalar
function in the same regulator and boundary-condition class as
\(\partial_\mu A^\mu\).  Along the Abelian gauge orbit
\[
  A_\mu^\xi=A_\mu+\partial_\mu\xi,
  \qquad
  \psi^\xi=\ee^{\ii g\xi}\psi,
  \qquad
  \bar\psi^\xi=\bar\psi\,\ee^{-\ii g\xi},
\]
the linearized orbit-slice map is
\begin{equation}
  \mathcal M_{\rm QED}(x,y)
  =
  \left.
  \frac{\delta F_\varphi[A^\xi](x)}{\delta\xi(y)}
  \right|_{\xi=0}
  =
  \Box_x\delta^{(4)}(x-y).
  \label{eq:qed-fp-operator-box}
\end{equation}
After residual zero modes have been removed or separately fixed,
\[
  \Delta_{\rm FP}^{U(1)}=\det{}'\Box .
\]
This determinant is independent of \(A_\mu,\psi,\bar\psi\).  Hence it cancels
from normalized representative Green functions and may be absorbed into the
normalization of \(Z\).  If it is represented by anticommuting ghosts, the
ghost action is the free decoupled quadratic action
\[
  S_{\rm gh}^{U(1)}
  =
  \int\dd^4x\,\bar c\,\Box c ,
\]
so there is no QED ghost vertex in the perturbative rules below.

The gauge condition depends only on the vector-potential representative, so
the matter-field phases do not enter the Faddeev--Popov operator.  Directly,
\[
  F_\varphi[A^\xi](x)
  =
  \partial_\mu A^\mu(x)+\Box\xi(x)-\varphi(x).
\]
Taking the functional derivative with respect to \(\xi(y)\) gives
\eqref{eq:qed-fp-operator-box}.  Since the result contains no field, the
Jacobian is a constant determinant on the regular gauge-parameter subspace.
The prime on \(\det{}'\Box\) records the same zero-mode separation already
needed for the gauge slice itself: constant gauge parameters, or any residual
modes allowed by the chosen boundary conditions, do not change
\(\partial_\mu A^\mu\).  For normalized correlators the same determinant
multiplies numerator and denominator.  The Grassmann representation follows
from the finite-dimensional Berezin identity
\(\int D\bar c\,Dc\,\exp(\ii\bar cMc)\propto\det M\), applied to
\(M=\Box\).

This is the Abelian, linear-gauge specialization of the general
local orbit-measure construction in
Proposition~\ref{prop:fp-orbit-measure-construction} and
\eqref{eq:fp-gauge-fixed-integral}.  In a nonabelian gauge theory the orbit
map is instead
\[
  \delta_\zeta A_\mu^a=(D_\mu(A)\zeta)^a,
  \qquad
  \mathcal M^{ab}(A;x,y)
  =
  \partial_x^\mu D_\mu^{ab}(A;x)\,\delta^{(4)}(x-y)
\]
for Lorenz gauge.  The determinant then depends on the gauge representative
and its ghost representation produces interacting ghost vertices.  Moreover,
outside a sufficiently small regular chart the Faddeev--Popov operator can
develop zero modes and a single gauge condition can meet an orbit more than
once.  Those issues are not QED Feynman-rule details; they belong to the
nonabelian local-coordinate and BRST analysis of
Chapter~\ref{ch:volume-ii-18}, especially
Quoted Theorem~\ref{thm:singer-gribov-obstruction}.

\begin{definition}[Regulated gauge-fixed QED source functional]
\label{def:qed-gauge-fixed-source-functional}
Introduce an ordinary source \(J^\mu\) for \(A_\mu\), and Grassmann sources
\(\eta_\alpha,\bar\eta^\alpha\) for \(\bar\psi^\alpha,\psi_\alpha\). The
regulated Lorentzian generating functional, written in continuum shorthand, is
\[
\begin{aligned}
  Z[J,\bar\eta,\eta]
  &=
  \int DA\,D\bar\psi\,D\psi\,
  \exp\Biggl\{
    \ii\int\dd^4x\,
    \bigl(\mathcal L_{\mathrm{QED}}+\Delta\mathcal L\bigr) \\
  &\hspace{7em}
    +\ii\int\dd^4x\,
    \bigl(J^\mu A_\mu+\bar\eta\psi+\bar\psi\eta\bigr)
  \Biggr\}.
\end{aligned}
\]
As in the preceding finite-dimensional Berezin discussion, this notation is
shorthand for a regulated Gaussian integral followed by perturbative
expansion of the interaction term
\[
  \ii gA_\mu\bar\psi\gamma^\mu\psi
\]
around the Gaussian Maxwell and Dirac functionals already constructed.  The
covariant gauge parameter \(\xi_{\mathrm g}\) changes representative
correlators; the local gauge-invariant correlators obtained from neutral
operators and field strengths do not depend on this representative choice.
\end{definition}

\paragraph{Representative QED Feynman rules.}
\label{par:qed-representative-feynman-rules}
The coefficientwise expansion of
Definition~\ref{def:qed-gauge-fixed-source-functional} gives the following
momentum-space representative rules. A directed fermion line from
\(\psi_\alpha\) to \(\bar\psi^\beta\) carries
\[
\begin{tikzpicture}[baseline=-0.6ex,scale=0.82]
  \draw[thick,-{Stealth[length=3pt]}] (0,0) -- (1.8,0);
  \node[above] at (0.9,0) {\scriptsize \(k\)};
  \node[below] at (0,0) {\scriptsize \(\psi_\alpha\)};
  \node[below] at (1.8,0) {\scriptsize \(\bar\psi^\beta\)};
\end{tikzpicture}
\quad
=
\frac{-\ii(-\ii\not k+m)_\alpha{}^\beta}
     {k^2+m^2-\ii\epsilon}.
\]
The photon representative propagator is
\[
\begin{tikzpicture}[baseline=-0.6ex,scale=0.82]
  \draw[thick,decorate,decoration={snake,amplitude=1.0pt,segment length=5pt}]
    (0,0) -- (1.8,0);
  \node[above] at (0.9,0) {\scriptsize \(k\)};
  \node[below] at (0,0) {\scriptsize \(A_\mu\)};
  \node[below] at (1.8,0) {\scriptsize \(A_\nu\)};
\end{tikzpicture}
\quad
=
\frac{-\ii}{k^2-\ii\epsilon}
\left[
  \eta_{\mu\nu}
  -(1-\xi_{\mathrm g})\frac{k_\mu k_\nu}{k^2-\ii\epsilon}
\right].
\]
The cubic vertex with one photon representative and two spinor
representatives is
\[
\begin{tikzpicture}[baseline=-0.5ex,scale=0.82]
  \coordinate (v) at (0,0);
  \draw[thick,decorate,decoration={snake,amplitude=1.0pt,segment length=5pt}]
    (-1.0,0) -- (v);
  \draw[thick,-{Stealth[length=3pt]}] (v) -- (0.85,0.62);
  \draw[thick,-{Stealth[length=3pt]}] (0.85,-0.62) -- (v);
  \fill (v) circle (1.7pt);
  \node[left] at (-1.0,0) {\scriptsize \(\mu\)};
  \node[right] at (0.85,0.62) {\scriptsize \(\beta\)};
  \node[right] at (0.85,-0.62) {\scriptsize \(\alpha\)};
\end{tikzpicture}
\quad
=
-g(\gamma^\mu)_\beta{}^\alpha .
\]
For the electron convention \(g=-e\), the vertex factor is
\(e(\gamma^\mu)_\beta{}^\alpha\). The gauge parameter
\(\xi_{\mathrm g}\) appears only in the representative photon propagator.
Amplitudes between physical photon helicity states are obtained after LSZ
projection and are independent of the longitudinal representative.

The fermion propagator is the inverse of the free Dirac kernel in the
conventions of Chapter~\ref{ch:spinor-fields-fermionic-path-integrals}; the
displayed numerator is the adjugate of \(\not k+\ii m\) in the mostly-plus
spinor convention, with the Feynman boundary value fixed by
\(-\ii\epsilon\).  The photon representative propagator is the inverse of the
quadratic Maxwell operator after adding
\(-(\partial_\mu A^\mu)^2/(2\xi_{\mathrm g})\), as derived in
Chapter~\ref{ch:maxwell-theory-constraints-gauge-fixing}.  The Abelian
Faddeev--Popov determinant is field independent by the Abelian specialization
above, so no ghost vertex occurs.  Finally, differentiating the interaction
\(\ii\int \ii g A_\mu\bar\psi\gamma^\mu\psi\) with respect to one photon
source and the two Grassmann sources gives the cubic representative vertex
\(-g\gamma^\mu\), with the displayed index placement inherited from the
ordered source convention in
Definition~\ref{def:qed-gauge-fixed-source-functional}.

\section{External Factors as Pole Data}

\begin{definition}[QED external pole-residue factors]
\label{def:qed-external-pole-residue-factors}
External factors in a diagrammatic amplitude are residues of poles already
identified by LSZ. The particle definitions have already been supplied by the
one-particle spectral data and scattering construction. With the normalizations
of Chapters~\ref{ch:lsz-reduction},
\ref{ch:spinor-fields-fermionic-path-integrals}, and
\ref{ch:maxwell-theory-constraints-gauge-fixing}, the factors are
\[
\begin{array}{c|c}
\text{external state} & \text{LSZ factor} \\ \hline
e^-_{\mathrm{out}}(\vec p,\sigma)
&
\displaystyle
\frac{Z_\psi^{1/2}}{(2\pi)^{3/2}}\,
(\bar u^\sigma)^\alpha(\vec p)
\\[1.1em]
e^+_{\mathrm{out}}(\vec p,\sigma)
&
\displaystyle
\frac{Z_\psi^{1/2}}{(2\pi)^{3/2}}\,
v^\sigma_\alpha(\vec p)
\\[1.1em]
\gamma_{\mathrm{out}}(\vec k,h)
&
\displaystyle
\frac{Z_A^{1/2}}{(2\pi)^{3/2}\sqrt{2|\vec k|}}\,
e_\mu^{h*}(\vec k)
\\[1.1em]
e^-_{\mathrm{in}}(\vec p,\sigma)
&
\displaystyle
\frac{Z_\psi^{1/2}}{(2\pi)^{3/2}}\,
u^\sigma_\alpha(\vec p)
\\[1.1em]
e^+_{\mathrm{in}}(\vec p,\sigma)
&
\displaystyle
\frac{Z_\psi^{1/2}}{(2\pi)^{3/2}}\,
(\bar v^\sigma)^\alpha(\vec p)
\\[1.1em]
\gamma_{\mathrm{in}}(\vec k,h)
&
\displaystyle
\frac{Z_A^{1/2}}{(2\pi)^{3/2}\sqrt{2|\vec k|}}\,
e_\mu^h(\vec k).
\end{array}
\]
\end{definition}

\begin{remark}[Status of the charged entries]
The electron and positron rows are literal LSZ residues only in a sector whose
charged spectrum contains an isolated massive one-particle shell.  This is the
case for an infrared-deformed theory, for instance a theory with a small
gauge-invariant photon mass regulator, and it is the pole bookkeeping used in
the hard part of perturbation theory.  It is not an exact statement about
massless four-dimensional QED after the infrared regulator is removed.  Gauss'
law ties nonzero electric charge to an electromagnetic field at spatial
infinity, while the massless photon allows arbitrarily soft excitations with
the same total charge and arbitrarily nearby energy-momentum.  The charged
threshold is therefore not separated from a continuum of charge plus soft
photon states, and the hypothesis needed by ordinary LSZ--a delta atom or
isolated pole associated with a normalizable one-electron Wigner subspace--is
absent.  The spinor factors below should consequently be read as hard-kernel
data whose physical completion is supplied by the inclusive and dressed
constructions of Chapter~\ref{ch:volume-ii-infrared-inclusive-qed}.
\end{remark}

\begin{proposition}[Residue origin of the external factors]
\label{prop:qed-external-factors-lsz-residue-origin}
Assume the spinor and photon representative two-point functions have the
isolated one-particle pole data specified in Chapters~\ref{ch:lsz-reduction},
\ref{ch:spinor-fields-fermionic-path-integrals}, and
\ref{ch:maxwell-theory-constraints-gauge-fixing}.  Then the factors in
Definition~\ref{def:qed-external-pole-residue-factors} are exactly the
coefficients obtained by applying the LSZ boundary-value operation to the
corresponding amputated kernel.  A change
\[
  e_\mu^h(\vec k)\mapsto e_\mu^h(\vec k)+\alpha k_\mu
\]
does not change the physical photon state; it changes only the representative
used before the final helicity quotient.
\end{proposition}

\begin{proof}
The spinor entries use the normalization derived in
\eqref{eq:spinor-external-normalization-factors}: the \(2\omega_{\vec p}\)
factor in the covariant one-particle measure is paired with the
\(1/(2\omega_{\vec p})\) in the spinor completeness relation, leaving the
Fourier factor \((2\pi)^{-3/2}\) and the pole residue \(Z_\psi^{1/2}\).
The free spinor or photon line adjacent to the external leg is not retained:
the LSZ operator removes that propagator and evaluates the residue at the
one-particle pole. The displayed spinor index is contracted with the
corresponding open index of the amputated Green-function kernel. In a
Fourier convention in which all momenta flow into a Green function, outgoing
massive particles are represented by the boundary value with the opposite
energy sign; the table above is written directly in terms of the physical
future-directed external momenta.

The massive spinor momenta satisfy
\[
  p^0=\omega_{\vec p}=\sqrt{\vec p^{\,2}+m^2},
\]
and the photon momentum satisfies
\[
  k^0=|\vec k|,\qquad k^2=0 .
\]
The spinors obey
\[
  (\not p-\ii m)u^\sigma(\vec p)=0,\qquad
  (\not p+\ii m)v^\sigma(\vec p)=0,
\]
with the corresponding adjoint equations for \(\bar u,\bar v\). The photon
polarization representatives obey
\[
  k^\mu e_\mu^h(\vec k)=0,\qquad
  e_\mu^h(\vec k)\sim e_\mu^h(\vec k)+\alpha k_\mu .
\]
For any polarization representative,
\[
  \not e^{\,h}(\vec k)=e_\mu^h(\vec k)\gamma^\mu .
\]
The LSZ operation removes the adjacent free propagator and leaves precisely
the displayed residue vector or covector.  For the photon, the quotient by
longitudinal representatives is the helicity quotient constructed in
Chapter~\ref{ch:massless-particles-helicity-gauge-redundancy}; replacing
\(e_\mu^h\) by \(e_\mu^h+\alpha k_\mu\) changes the representative but not
the resulting one-photon helicity state.
\end{proof}

\begin{center}
\begin{tikzpicture}[>=Stealth,x=1cm,y=1cm]
  \tikzset{
    flowbox/.style={
      draw,
      rounded corners=2pt,
      align=center,
      font=\scriptsize,
      minimum height=0.72cm,
      text width=2.25cm
    },
    flowlabel/.style={font=\scriptsize,fill=white,inner sep=1.2pt}
  }
  \node[flowbox] (green) at (0,0) {gauge-fixed\\Green functions};
  \node[flowbox] (poles) at (3.35,0) {one-particle\\poles};
  \node[flowbox] (hard) at (6.7,0) {hard\\amplitude};
  \draw[->,thick] (green) -- node[flowlabel,above] {LSZ} (poles);
  \draw[->,thick] (poles) -- node[flowlabel,above] {residues} (hard);
  \draw[red!70!black,thick] (8.25,-0.72) -- (8.25,0.72);
  \node[red!70!black,align=center] at (9.35,0)
    {\scriptsize infrared\\[-0.1em]\scriptsize completion};
\end{tikzpicture}
\end{center}

The vertical mark in the diagram indicates a boundary of this chapter. The
hard-amplitude rules are perturbative rules for fixed external momenta and
helicities. The infrared completion of charged scattering is a further
construction.

Diagrammatically, the external factors attach to amputated kernels as follows:
\[
\begin{array}{cc}
\begin{tikzpicture}[baseline=(current bounding box.center),>=Stealth,scale=0.86]
  \tikzset{
    fermion/.style={line width=0.55pt,-{Stealth[length=3pt]}},
    blob/.style={circle,draw,fill=gray!16,minimum size=0.72cm,inner sep=0pt}
  }
  \node[blob] (b) at (1.25,0) {};
  \draw[fermion] (0,0)--(b);
  \node[below] at (0.62,-0.05) {\scriptsize \(u_\alpha^\sigma(\vec p)\)};
\end{tikzpicture}
&
\begin{tikzpicture}[baseline=(current bounding box.center),>=Stealth,scale=0.86]
  \tikzset{
    fermion/.style={line width=0.55pt,-{Stealth[length=3pt]}},
    blob/.style={circle,draw,fill=gray!16,minimum size=0.72cm,inner sep=0pt}
  }
  \node[blob] (b) at (0,0) {};
  \draw[fermion] (b)--(1.25,0);
  \node[below] at (0.85,-0.05) {\scriptsize \((\bar u^\sigma)^\alpha(\vec p)\)};
\end{tikzpicture}
\\[1.5em]
\begin{tikzpicture}[baseline=(current bounding box.center),>=Stealth,scale=0.86]
  \tikzset{
    photon/.style={decorate,decoration={snake,amplitude=1.1pt,segment length=5pt},line width=0.55pt},
    blob/.style={circle,draw,fill=gray!16,minimum size=0.72cm,inner sep=0pt}
  }
  \node[blob] (b) at (1.25,0) {};
  \draw[photon] (0,0)--(b);
  \node[below] at (0.55,-0.05) {\scriptsize \(e_\mu^h(\vec k)\)};
\end{tikzpicture}
&
\begin{tikzpicture}[baseline=(current bounding box.center),>=Stealth,scale=0.86]
  \tikzset{
    photon/.style={decorate,decoration={snake,amplitude=1.1pt,segment length=5pt},line width=0.55pt},
    blob/.style={circle,draw,fill=gray!16,minimum size=0.72cm,inner sep=0pt}
  }
  \node[blob] (b) at (0,0) {};
  \draw[photon] (b)--(1.25,0);
  \node[below] at (0.82,-0.05) {\scriptsize \(e_\mu^{h*}(\vec k)\)};
\end{tikzpicture}
\end{array}
\]
with the common normalization factors displayed in the table.

\section{Compton Scattering at Tree Level}

\paragraph{Tree-level Compton hard kernel.}
\label{par:qed-tree-compton-hard-kernel}
Consider the process
\[
  e^-(p,\sigma)+\gamma(k,h)
  \longrightarrow
  e^-(p',\sigma')+\gamma(k',h')
\]
with
\[
  p+k=p'+k' .
\]
At tree level two diagrams contribute:
\[
\begin{tikzpicture}[scale=0.93,baseline=(current bounding box.center)]
  \tikzset{
    photon/.style={decorate,decoration={snake,amplitude=1.1pt,segment length=5pt},line width=0.58pt}
  }
  \begin{scope}
    \draw[thick,-{Stealth[length=3pt]}] (0,0) -- (1.25,0) -- (2.5,0);
    \draw[photon,-{Stealth[length=3pt]}] (0.55,1.0) -- (0.98,0.08);
    \draw[photon,-{Stealth[length=3pt]}] (1.52,0.08) -- (1.95,1.0);
    \node[below] at (0,0) {\scriptsize \(p\)};
    \node[below] at (1.25,0) {\scriptsize \(p+k\)};
    \node[below] at (2.5,0) {\scriptsize \(p'\)};
    \node[above] at (0.52,1.0) {\scriptsize \(k\)};
    \node[above] at (2.0,1.0) {\scriptsize \(k'\)};
  \end{scope}
  \node at (3.05,0.22) {\Large \(+\)};
  \begin{scope}[xshift=3.75cm]
    \draw[thick,-{Stealth[length=3pt]}] (0,0) -- (1.25,0) -- (2.5,0);
    \draw[photon,-{Stealth[length=3pt]}] (0.55,1.0) -- (0.98,0.08);
    \draw[photon,-{Stealth[length=3pt]}] (1.52,0.08) -- (1.95,1.0);
    \node[below] at (0,0) {\scriptsize \(p\)};
    \node[below] at (1.25,0) {\scriptsize \(p-k'\)};
    \node[below] at (2.5,0) {\scriptsize \(p'\)};
    \node[above] at (0.52,1.0) {\scriptsize \(k'\)};
    \node[above] at (2.0,1.0) {\scriptsize \(k\)};
  \end{scope}
\end{tikzpicture}
\]
After removing the factor
\[
  (2\pi)^4\delta^{(4)}(p+k-p'-k')
\]
from the scattering matrix element, the reduced amplitude is
\[
\begin{aligned}
  \ii\mathcal M
  &=
  \frac{-\ii e^2}
       {(2\pi)^6\sqrt{2|\vec k|}\sqrt{2|\vec k'|}}\,
  \bar u^{\sigma'}(\vec p')
  \Biggl[
    \not e^{\,h'*}(\vec k')\,
    \frac{-\ii(\not p+\not k)+m}
         {(p+k)^2+m^2-\ii\epsilon}\,
    \not e^{\,h}(\vec k)  \\
  &\hspace{8.7em}
    +
    \not e^{\,h}(\vec k)\,
    \frac{-\ii(\not p-\not k')+m}
         {(p-k')^2+m^2-\ii\epsilon}\,
    \not e^{\,h'*}(\vec k')
  \Biggr]
  u^\sigma(\vec p).
\end{aligned}
\]

At tree level the source expansion of
the representative QED Feynman rules above has exactly two
connected diagrams with one internal electron line: the channel with internal
momentum \(p+k\) and the channel with internal momentum \(p-k'\).  The photon
external factors are the helicity representatives in
Definition~\ref{def:qed-external-pole-residue-factors}; the adjacent photon
propagators have already been removed by LSZ.  The two cubic vertices
contribute the factor \(e^2\) in the electron convention \(g=-e\), while the
single internal spinor propagator contributes the displayed numerator and
denominator in each channel.  Removing the overall
\((2\pi)^4\delta^{(4)}(p+k-p'-k')\) gives the stated reduced kernel.

In an infrared-deformed theory with ordinary charged one-particle poles, this
is an LSZ-reduced expression: the external spinors and polarization vectors
are the pole residues listed above.  In massless QED it is the Born hard
kernel.  It becomes part of a physical charged-scattering prediction only
after the infrared completion has been specified.  At higher orders, virtual
soft photons and real photons below detector resolution cannot be separated
into an exact fixed-photon-number electron \(S\)-matrix element; they are
combined in the inclusive or dressed constructions developed in
Chapter~\ref{ch:volume-ii-infrared-inclusive-qed}.

\begin{theorem}[Tree-level Ward identity for the Compton hard kernel]
\label{thm:qed-compton-tree-ward-identity}
The tree-level Compton hard kernel displayed above
descends from photon polarization representatives to photon helicity states.
Equivalently, replacing either external photon polarization representative by
that representative plus a multiple of its null momentum leaves the sum of the
two tree diagrams unchanged.
\end{theorem}

\begin{proof}
The expression descends from photon representatives to photon helicity states.
Let
\[
  N(q)=-\ii\not q+m,\qquad D(q)=q^2+m^2-\ii\epsilon .
\]
If the incoming polarization is shifted by
\[
  e_\mu^h(k)\mapsto e_\mu^h(k)+\alpha k_\mu ,
\]
then the first diagram changes by a factor proportional to
\[
  \frac{N(p+k)\not k}{D(p+k)}u^\sigma(p).
\]
Using \(p^2=-m^2\), \(k^2=0\), and
\((\not p-\ii m)u^\sigma(p)=0\), one obtains
\[
  N(p+k)\not k\,u^\sigma(p)
  =
  -2\ii(p\cdot k)u^\sigma(p),
  \qquad
  D(p+k)=2p\cdot k-\ii\epsilon .
\]
Thus the first diagram contributes
\[
  -\ii\alpha\,
  \bar u^{\sigma'}(p')\not e^{\,h'*}(k')u^\sigma(p)
\]
up to the common prefactor. For the second diagram, momentum conservation
gives \(p-k'=p'-k\), and the adjoint Dirac equation gives
\[
  \bar u^{\sigma'}(p')\not k\,N(p-k')
  =
  -2\ii(p'\cdot k)\bar u^{\sigma'}(p'),
  \qquad
  D(p-k')=-2p'\cdot k-\ii\epsilon .
\]
The second diagram therefore contributes
\[
  +\ii\alpha\,
  \bar u^{\sigma'}(p')\not e^{\,h'*}(k')u^\sigma(p)
\]
with the same common prefactor. The two incoming-polarization shifts cancel.

Now shift the outgoing representative by
\[
  e_\mu^{h'*}(k')\mapsto e_\mu^{h'*}(k')+\beta k'_\mu .
\]
In the first diagram, \(p+k=p'+k'\), and
\[
  \bar u^{\sigma'}(p')\not k'\,N(p+k)
  =
  -2\ii(p'\cdot k')\bar u^{\sigma'}(p'),
  \qquad
  D(p+k)=2p'\cdot k'-\ii\epsilon .
\]
It contributes
\[
  -\ii\beta\,
  \bar u^{\sigma'}(p')\not e^{\,h}(k)u^\sigma(p)
\]
up to the common prefactor. In the second diagram,
\[
  N(p-k')\not k'\,u^\sigma(p)
  =
  -2\ii(p\cdot k')u^\sigma(p),
  \qquad
  D(p-k')=-2p\cdot k'-\ii\epsilon ,
\]
so the contribution is
\[
  +\ii\beta\,
  \bar u^{\sigma'}(p')\not e^{\,h}(k)u^\sigma(p).
\]
The outgoing-polarization shifts cancel as well. This cancellation is the
tree-level Ward identity for the Compton amplitude.
\end{proof}

With the cross-section normalization of
Chapter~\ref{ch:cross-sections-unitarity}, the total cross section
for fixed incoming spin and helicity is
\[
\begin{aligned}
  \sigma
  &=
  \frac{(2\pi)^6}{v}
  \int\dd^3\vec k'\,\dd^3\vec p'\,
  (2\pi)^4\delta^{(4)}(k+p-k'-p')  \\
  &\qquad\times
  \sum_{\substack{h'=\pm1\\ \sigma'=1,2}}
  \left|
    \mathcal M(\vec p',\sigma';\vec k',h'
    \mid
    \vec p,\sigma;\vec k,h)
  \right|^2 ,
\end{aligned}
\]
where \(v\) is the incoming relative velocity. The spin and polarization sums
are obtained from the completeness relations of
Chapters~\ref{ch:spinor-fields-fermionic-path-integrals} and
\ref{ch:massless-particles-helicity-gauge-redundancy}.
The reduced Compton kernel \(\mathcal M\) is the connected on-shell
matrix element after removal of the total momentum-conservation delta
distribution.  In the sharp-kernel event-probability formula, the final
electron and photon momenta are integrated with the normalization used in the
external-state chapters, while the remaining
\((2\pi)^4\delta^{(4)}(k+p-k'-p')\) imposes energy-momentum conservation on
the event.  Summing over \(h'\) and \(\sigma'\) implements the inclusive
measurement over final physical helicity and spin labels.  The incoming spin
and helicity are specified rather than averaged, so no factor averaging over
initial labels appears.

\section{Infrared Boundary of Charged Scattering}

\begin{remark}[Infrared boundary of charged QED scattering]
The perturbative rules above are rules for hard Green-function residues. They
are valid order by order once an infrared regulator or an inclusive
prescription has been specified. The exact long-range theory has an additional
constraint: Gauss' law ties electric charge to an electromagnetic field
extending to arbitrarily large distances. Consequently the charged-sector
asymptotic data is infrared-sensitive: its spectral support is tied to the
charged mass threshold together with arbitrarily soft radiation. This lies
outside the isolated massive-particle hypothesis used in the basic
Haag--Ruelle construction.

This boundary is already visible in the charged dressing
\(\Psi_{C_x}(x)\). A dressing that retains nonzero electric charge extends to
infinity. In scattering, the same long-range field is represented
perturbatively by arbitrarily soft photons.

For a gauge parameter that approaches a nonzero constant on a large sphere,
Gauss' law relates the total electric charge inside the sphere to the
asymptotic electric flux.  Thus a charged vector cannot be created from the
vacuum by a compactly supported gauge-invariant operator, and a charged
insertion must carry dressing data extending to the boundary as in
Definition~\ref{def:qed-wilson-line-dressed-charged-insertion}.  Because the
photon is massless, adding arbitrarily low-energy photons to a charged hard
configuration changes the state while changing the total energy-momentum by
an arbitrarily small amount.  The charged mass threshold is therefore not an
isolated one-particle mass shell separated by a gap from the rest of the
charged spectrum.  This is precisely the hypothesis ordinary massive
Haag--Ruelle and the pole form of LSZ require, so the representative hard
kernel must be completed by an infrared-aware construction.
\end{remark}

The physical completions are inclusive transition probabilities, where soft
photons below a resolution threshold are summed over, or dressed asymptotic
charged states. Those constructions belong after renormalization and the
analysis of infrared singularities. The present chapter supplies the
gauge-fixed hard-amplitude rules on which that later analysis acts.
