\chapter{Cross Sections, Phase Space, and Unitarity}
\label{ch:cross-sections-unitarity}
\ConventionsLorentzian


\section{Transition Probabilities}

\begin{definition}[Transition probability for an outgoing event class]
\label{def:scattering-transition-probability-event-class}
Let \(\mathcal H_{\rm as}\) be the asymptotic Hilbert space in a sector where
the incoming and outgoing Haag--Ruelle ranges coincide.  The scattering
operator
\[
  S:\mathcal H_{\rm as}\to\mathcal H_{\rm as}
\]
is then unitary.  Write
\[
  S=\mathbf 1+\ii T .
\]
For a normalized incoming vector \(\Psi_{\mathrm{in}}\in\mathcal H_{\rm as}\) and
an outgoing class of states \(\mathcal D\), let
\(\Pi_{\mathcal D}^{\mathrm{out}}\) be the orthogonal projection onto the closed
subspace spanned by \(\mathcal D\).  The transition probability into
\(\mathcal D\) is
\[
  P_{\Psi\to\mathcal D}
  =
  \left\|
    \Pi_{\mathcal D}^{\mathrm{out}}S\Psi_{\mathrm{in}}
  \right\|^2 .
\]
This is the primary definition.  Sharp-momentum formulas are distributional
kernels obtained from this expression after smearing with wave packets and
then taking a narrow-packet limit.
\end{definition}

\begin{definition}[Invariant amplitude convention]
\label{def:invariant-amplitude-convention}
For positive-energy momenta in the mostly-plus metric,
\[
  p_i=(E_i,\vec p_i),
  \qquad
  q_a=(E_a,\vec q_a),
\]
with
\[
  p_i^2=-m_i^2,\qquad q_a^2=-M_a^2,
\]
set
\[
  P_{\mathrm{in}}=\sum_i p_i,
  \qquad
  P_{\mathrm{out}}=\sum_a q_a .
\]
The invariant amplitude convention inherited from LSZ is
\[
  \langle q_1,\ldots,q_m|T|p_1,\ldots,p_n\rangle
  =
  (2\pi)^D\delta^{(D)}(P_{\mathrm{out}}-P_{\mathrm{in}})
  \mathcal M(q_1,\ldots,q_m|p_1,\ldots,p_n).
\]
The bra and ket are generalized basis vectors in \(\mathcal H_{\rm as}\).
They are elements of the appropriate dual test-vector space, as in
Section~\ref{sec:rigged-hilbert-continuous-spectrum}. Equivalently, the same
distribution is obtained from the connected part of the Haag--Ruelle kernel
\[
  {}_{\mathrm{out}}\langle q_1,\ldots,q_m|p_1,\ldots,p_n\rangle_{\mathrm{in}}
\]
after subtracting the identity contribution.  The connected contribution to
\(S\) is therefore
\[
  \ii(2\pi)^D\delta^{(D)}(P_{\mathrm{out}}-P_{\mathrm{in}})\mathcal M .
\]
Throughout this chapter, \(\mathcal M(f|i)\) means the \(T\)-matrix element
with ket label \(i\) and bra label \(f\), namely the transition \(i\to f\).
The notation is ordered; \(\mathcal M(f|i)\) and \(\mathcal M(i|f)\) are
different matrix elements unless an additional reciprocity symmetry is
assumed.
\end{definition}

\section{Sharp-Momentum Limit and Delta-Squares}

\begin{proposition}[Sharp-momentum density limit]
\label{prop:sharp-momentum-density-limit-cross-section}
The wave-packet definition is primary, but the conventional plane-wave formula
is useful as a check on all factors.  Temporarily use
\(\delta\)-normalized one-particle states in \(d=D-1\) spatial dimensions,
\[
  {}_\delta\langle \vec q\,|\vec p\rangle_\delta
  =
  \delta^{(d)}(\vec q-\vec p),
\]
and write the connected kernel as
\[
  S^{\rm conn}_{\beta\alpha}
  =
  \ii(2\pi)^D\delta^{(D)}(P_\beta-P_\alpha)M_\delta(\beta|\alpha).
\]
The subscript in \(M_\delta\) is part of the notation: it records the
\(\delta^{(d)}\)-normalized sharp-momentum basis used in this temporary
calculation.  It is neither the invariant amplitude \(\mathcal M\) nor a
partial-wave coefficient.  Later in this chapter the invariant amplitude is
expanded in partial waves with coefficients \(a_\ell(s)\), while \(S_\ell(s)\)
denotes an eigenvalue of the scattering operator on a partial-wave channel:
\[
\begin{array}{c|l}
  M_\delta(\beta|\alpha)
    & \delta^{(d)}\text{-normalized connected sharp-momentum kernel},\\
  \mathcal M(f|i)
    & \text{invariant amplitude after external normalization factors},\\
  a_\ell(s)
    & \text{coefficient in the ordered }16\pi\text{ partial-wave expansion},\\
  S_\ell(s)
    & \text{elastic partial-wave }S\text{-matrix eigenvalue}.
\end{array}
\]
Let \(\mathcal D\) be a final region that does not contain the identity
no-scattering contribution.  Squaring the distributional kernel gives
\[
  P_{\alpha\to\mathcal D}
  =
  (2\pi)^D\delta^{(D)}(0)
  \int_{\mathcal D}\dd\beta_\delta\,
  (2\pi)^D\delta^{(D)}(P_\beta-P_\alpha)
  |M_\delta(\beta|\alpha)|^2 ,
\]
where \(\dd\beta_\delta\) is the product of the final \(\dd^d\vec q\) measures,
with the resolution-of-identity factors appropriate to the outgoing Fock
sector.  A finite box
regularization of spatial volume \(V\) and observation time \(\mathcal T\) gives
\[
  (2\pi)^D\delta^{(D)}(0)=V\mathcal T,
  \qquad
  \delta^{(d)}(0)=\frac{V}{(2\pi)^d}.
\]
For two sharp incoming particles, the normalized box state carries the factor
\(\delta^{(d)}(0)^{-2}\).  Hence
\[
  P_{\alpha\to\mathcal D}
  =
  \frac{(2\pi)^{2d}\mathcal T}{V}
  \int_{\mathcal D}\dd\beta_\delta\,
  (2\pi)^D\delta^{(D)}(P_\beta-P_\alpha)
  |M_\delta(\beta|\alpha)|^2 .
\]
If the two beams have relative speed \(v_{\rm rel}=|\vec v_1-\vec v_2|\) in
the frame used for the box, the geometric probability to scatter into
\(\mathcal D\) during time \(\mathcal T\) is
\[
  P_{\alpha\to\mathcal D}
  =
  \frac{\sigma_{\mathcal D}v_{\rm rel}\mathcal T}{V}.
\]
Therefore
\[
  \sigma_{\mathcal D}
  =
  \frac{(2\pi)^{2D-2}}{v_{\rm rel}}
  \int_{\mathcal D}\dd\beta_\delta\,
  (2\pi)^D\delta^{(D)}(P_\beta-P_\alpha)
  |M_\delta(\beta|\alpha)|^2 .
\]
Converting \(M_\delta\) to the invariant amplitude \(\mathcal M\) by the
external normalization factors stated in the LSZ chapter gives the covariant
phase-space formula below.
\end{proposition}

\begin{proof}
The connected sharp kernel contains a single total momentum-conservation
distribution.  In the probability, the square of this distribution is defined
by putting the theory in a spatial box of volume \(V\) and evolving for an
observation time \(\mathcal T\).  In that regulator,
\[
  \bigl[(2\pi)^D\delta^{(D)}(P_\beta-P_\alpha)\bigr]^2
  =
  (2\pi)^D\delta^{(D)}(0)\,
  (2\pi)^D\delta^{(D)}(P_\beta-P_\alpha)
\]
with \((2\pi)^D\delta^{(D)}(0)=V\mathcal T\).  A two-particle incoming
\(\delta\)-normalized box state contributes
\(\delta^{(d)}(0)^{-2}=[(2\pi)^d/V]^2\).  Multiplying these factors gives the
displayed probability density
\((2\pi)^{2d}\mathcal T/V\) times the final-state phase-space integral.
Dividing by the geometric probability
\(v_{\rm rel}\mathcal T/V\) for a unit transverse area gives the sharp
cross-section density.  The last conversion is exactly the external
normalization conversion between \(M_\delta\) and \(\mathcal M\) fixed in
Definition~\ref{def:invariant-amplitude-convention}.
\end{proof}

\section{Wave-Packet Origin of Cross Sections}

Cross sections are transition probabilities per incoming transverse flux.
Let \(\psi_1,\psi_2\in\Hilb_1\) be normalized incoming one-particle packets
with narrow momentum supports and distinct group velocities.  In the
center-of-mass frame let \(\vec p_*\) be the central momentum of the first
packet, and let
\[
  \mathbb R^{D-2}_\perp
  =
  \{b\in\mathbb R^{D-1}: b\cdot \vec p_*=0\}
\]
be the transverse impact-parameter plane.  If \(\tau_b\) denotes the spatial
translation acting on the one-particle Hilbert space, define
\[
  \Psi_b=\psi_1\odot \tau_b\psi_2 .
\]
Here \(\odot\) denotes the two-particle product in the asymptotic Fock space,
with the species labels and Bose or Fermi statistics appropriate to the
sector under consideration.  For an outgoing channel \(\mathcal D\) whose
projection removes the identity no-scattering component, the wave-packet
cross section is
\[
  \sigma_{\psi_1,\psi_2}(\mathcal D)
  =
  \int_{\mathbb R^{D-2}_\perp}\dd^{D-2}b\,
  \left\|
    \Pi_{\mathcal D}^{\mathrm{out}}S\Psi_b
  \right\|^2 .
\]
The dimension is length to the power \(D-2\), the dimension of transverse
area in \(D\) spacetime dimensions.

The plane-wave formula is the narrow-packet density of this Hilbert-space
quantity.  The Fourier transform in \(b\) enforces transverse momentum
conservation.  The long-time Haag--Ruelle localization enforces energy and
longitudinal momentum conservation in the scattering limit.  Together these
produce the single overall distribution
\[
  (2\pi)^D\delta^{(D)}(P_{\mathrm{out}}-P_{\mathrm{in}})
\]
appearing in the invariant amplitude convention.  The incoming current
through the transverse impact-parameter plane is the invariant relative speed
multiplied by the one-particle normalizations.  Thus the denominator in the
sharp-momentum two-particle formula is the invariant flux
\[
  \mathcal F(p_1,p_2)=4E_1E_2v_{\rm rel}.
\]
The next section expresses the same factor in covariant form.

\section{Invariant Phase Space and Flux}

\begin{definition}[Invariant phase space]
\label{def:invariant-final-state-phase-space}
For one on-shell final particle of mass \(M\), the Lorentz-invariant
positive-energy measure is
\[
  \dd\mu_M(q)
  =
  \frac{\dd^Dq}{(2\pi)^{D-1}}\,
  \theta(q^0)\delta(q^2+M^2)
  =
  \frac{\dd^{D-1}\vec q}{(2\pi)^{D-1}2E_{\vec q}},
\]
where
\[
  E_{\vec q}=\sqrt{\vec q^{\,2}+M^2}.
\]
For \(m\) final particles with total incoming momentum \(P\), define the
\(m\)-body phase-space measure
\[
  \dd\Phi_m(P;q_1,\ldots,q_m)
  =
  (2\pi)^D\delta^{(D)}\!\left(P-\sum_{a=1}^m q_a\right)
  \prod_{a=1}^m\dd\mu_{M_a}(q_a).
\]
This formula is written first for labeled outgoing particles.  The identical
particle factors come from the resolution of the identity on the asymptotic
Fock space.  For one bosonic species, the \(r\)-particle space is
\(\Hilb_1^{\odot r}\).  If
\[
  |\vec q_1,\ldots,\vec q_r\rangle_{\rm sym}
  =
  a^\dagger(\vec q_1)\cdots a^\dagger(\vec q_r)\Omega_F ,
\]
with the relativistic one-particle normalization used above, then the identity
on this symmetric sector is
\[
  \mathbf 1_{\Hilb_1^{\odot r}}
  =
  \frac1{r!}
  \int
  \prod_{j=1}^r\dd\mu_m(q_j)\,
  |\vec q_1,\ldots,\vec q_r\rangle_{\rm sym}
  {}_{\rm sym}\langle \vec q_1,\ldots,\vec q_r| .
\]
The factor \(1/r!\) compensates for the \(r!\) labeled tuples that represent
the same unordered Bose state.  For several species, the physical phase-space
measure carries the product
\[
  \prod_\lambda\frac1{n_\lambda!},
\]
where \(n_\lambda\) is the number of identical outgoing particles of species
\(\lambda\).  Fermionic sectors have the corresponding antisymmetric identity
resolution; the same factorial appears, with signs carried by the
antisymmetric external wavefunctions and matrix elements.
\end{definition}

\begin{proposition}[Invariant two-particle flux]
\label{prop:invariant-two-particle-flux}
For a two-particle initial state, the invariant flux factor is
\[
  \mathcal F(p_1,p_2)
  =
  4\sqrt{(p_1\cdot p_2)^2-m_1^2m_2^2}.
\]
Equivalently,
\[
  \mathcal F=4E_1E_2v_{\rm rel},
  \qquad
  v_{\rm rel}
  =
  \frac{\sqrt{(p_1\cdot p_2)^2-m_1^2m_2^2}}{E_1E_2}.
\]
In the center-of-mass frame,
\[
  p_1=(E_1,\vec p_*),
  \qquad
  p_2=(E_2,-\vec p_*),
\]
so
\[
  \mathcal F=4(E_1+E_2)|\vec p_*|.
\]
\end{proposition}

\begin{proof}
The quantity
\((p_1\cdot p_2)^2-m_1^2m_2^2\) is Lorentz invariant, so it suffices to
evaluate it in the center-of-mass frame.  With mostly-plus metric,
\[
  p_1\cdot p_2=-E_1E_2-\vec p_*^{\,2}.
\]
Using \(E_i^2=\vec p_*^{\,2}+m_i^2\),
\[
\begin{aligned}
  (p_1\cdot p_2)^2-m_1^2m_2^2
  &=
  (E_1E_2+\vec p_*^{\,2})^2
  -(E_1^2-\vec p_*^{\,2})(E_2^2-\vec p_*^{\,2})  \\
  &=
  \vec p_*^{\,2}(E_1+E_2)^2 .
\end{aligned}
\]
Thus \(\mathcal F=4(E_1+E_2)|\vec p_*|\).  Since
\(\vec v_1=\vec p_*/E_1\) and \(\vec v_2=-\vec p_*/E_2\), the relative speed
is
\[
  v_{\rm rel}=|\vec v_1-\vec v_2|
  =
  \frac{(E_1+E_2)|\vec p_*|}{E_1E_2},
\]
which proves the equivalent form \(\mathcal F=4E_1E_2v_{\rm rel}\).
\end{proof}

\begin{definition}[Differential cross section from invariant data]
\label{def:differential-cross-section-invariant-data}
The differential cross section for
\[
  p_1+p_2\longrightarrow q_1+\cdots+q_m
\]
is
\[
  \dd\sigma
  =
  \frac{1}{\mathcal F(p_1,p_2)}
  \left|\mathcal M(q_1,\ldots,q_m|p_1,p_2)\right|^2
  \dd\Phi_m(P_{\mathrm{in}};q_1,\ldots,q_m).
\]
All factors in this expression have now been fixed: the asymptotic state
normalization, the amplitude convention, the final-state measure, and the
incoming flux.
\end{definition}

\begin{center}
\begin{tikzpicture}[scale=0.86,>=Stealth]
  \draw[blue!75!black,very thick,->] (-4.4,-0.9)--(-1.35,-0.22);
  \draw[green!50!black,very thick,->] (-4.4,0.9)--(-1.35,0.22);
  \draw[purple!80!black,thick,fill=purple!7]
    (-1.15,0) ellipse (0.35 and 1.05);
  \draw[red!75!black,very thick,->] (-0.72,0.28)--(2.05,1.1);
  \draw[red!75!black,very thick,->] (-0.72,0)--(2.15,0.0);
  \draw[red!75!black,very thick,->] (-0.72,-0.28)--(2.0,-1.1);
  \node[blue!75!black] at (-3.45,-1.2) {$p_1,\ \vec v_1$};
  \node[green!50!black] at (-3.45,1.2) {$p_2,\ \vec v_2$};
  \node[purple!80!black,align=center] at (-1.15,-1.55)
    {effective\\area \(\sigma\)};
  \node[red!75!black] at (2.55,1.1) {$q_1$};
  \node[red!75!black] at (2.55,0.0) {$q_2$};
  \node[red!75!black] at (2.55,-1.1) {$q_m$};
  \node[align=center] at (4.1,0)
    {final\\channel};
  \node[align=center] at (-4.15,0)
    {\(\mathcal F=4E_1E_2v_{\rm rel}\)};
\end{tikzpicture}
\end{center}

\section{Two-Body Final States in Four Dimensions}

Set \(D=4\).  For a two-body final state, define
\[
  s=-(p_1+p_2)^2=P_{\mathrm{in}}^0{}^2-\vec P_{\mathrm{in}}^{\,2}.
\]
In the center-of-mass frame,
\[
  P_{\mathrm{in}}=(\sqrt s,\vec0).
\]
The K\"all\'en polynomial is
\[
  \lambda_K(s,a,b)=s^2+a^2+b^2-2sa-2sb-2ab .
\]
For incoming masses \(m_1,m_2\) and final masses \(M_1,M_2\),
\[
  |\vec p_*|
  =
  \frac{\sqrt{\lambda_K(s,m_1^2,m_2^2)}}{2\sqrt s},
  \qquad
  |\vec q_*|
  =
  \frac{\sqrt{\lambda_K(s,M_1^2,M_2^2)}}{2\sqrt s}.
\]
The next calculation performs the two-body phase-space integral over the
energy-conserving delta function.  Here \(\dd\Omega\) is the solid-angle
measure of one final momentum in the center-of-mass frame.

\paragraph{Two-body phase-space reduction in four dimensions.}
With the normalization of Definition~\ref{def:invariant-final-state-phase-space},
the two-body final-state measure in the center-of-mass frame is
\[
  \dd\Phi_2
  =
  \frac{1}{16\pi^2}
  \frac{|\vec q_*|}{\sqrt s}\,\dd\Omega .
\]
Consequently
\[
  \frac{\dd\sigma}{\dd\Omega}
  =
  \frac{1}{64\pi^2s}
  \frac{|\vec q_*|}{|\vec p_*|}
  |\mathcal M(s,\Omega)|^2 .
\]

In the center-of-mass frame the spatial delta function in
\(\dd\Phi_2\) sets \(\vec q_2=-\vec q_1\).  Writing
\(q=|\vec q_1|\), the remaining radial integral contains
\[
  \delta(\omega_1(q)+\omega_2(q)-\sqrt s),
  \qquad
  \omega_a(q)=\sqrt{q^2+M_a^2}.
\]
At the root \(q=|\vec q_*|\),
\[
  \frac{\dd}{\dd q}(\omega_1+\omega_2)
  =
  q\left(\frac1{\omega_1}+\frac1{\omega_2}\right)
  =
  \frac{q\sqrt s}{\omega_1\omega_2}.
\]
Therefore
\[
  \int\frac{q^2\dd q\,\dd\Omega}{(2\pi)^2\,4\omega_1\omega_2}\,
  \delta(\omega_1+\omega_2-\sqrt s)
  =
  \frac{|\vec q_*|}{16\pi^2\sqrt s}\,\dd\Omega .
\]
Dividing by the center-of-mass flux
\(\mathcal F=4\sqrt s\,|\vec p_*|\) gives the displayed differential
cross-section formula.

For two identical final particles, the symmetric two-particle identity
resolution supplies the factor \(1/2\) when the integral is written over the
full labeled solid angle.

For elastic scattering of identical scalar particles the preceding formula
already gives the total cross section.  Since
\(|\vec q_*|=|\vec p_*|\) and \(s=E_{\rm cm}^2\),
\[
  \sigma_{2\to2}^{\rm id}
  =
  \frac12
  \int\dd\Omega\,
  \frac{|\mathcal M(E_{\rm cm},\Omega)|^2}{64\pi^2s}
  =
  \frac{1}{2^7\pi^2E_{\rm cm}^2}
  \int\dd\Omega\,|\mathcal M(E_{\rm cm},\Omega)|^2 .
\]
The direct delta-function integration gives the same expression and is a useful
normalization check.  In the center-of-mass frame the spatial delta function
sets \(\vec q_2=-\vec q_1\).  Writing \(q=|\vec q_1|\), the radial
energy-conservation integral is
\[
  \int q^2\dd q\,
  \delta(\omega_1(q)+\omega_2(q)-E_{\rm cm})
  =
  \frac{q\,\omega_1\omega_2}{E_{\rm cm}}.
\]
Together with \(E_1E_2v_{\rm rel}=|\vec p_*|E_{\rm cm}\), this reduces the
labeled two-body phase-space integral to the same result above, with the
factor \(1/2\) supplied by the symmetric two-particle resolution of the
identity.

For the tree-level \(\phi^4\) example with a single scalar mass \(m\), the
elastic kinematics obey
\[
  |\vec q_*|=|\vec p_*|,
  \qquad
  \sqrt s=E_{\rm cm}.
\]
If the tree-level \(\phi^4\) invariant amplitude is
\[
  \mathcal M=-g+O(g^2),
\]
then
\[
  \sigma_{2\to2}
  =
  \frac{g^2}{32\pi s}+O(g^3)
\]
for identical final particles.

\section{Unitarity and the Optical Theorem}

\begin{theorem}[Sharp-kernel unitarity and the optical theorem]
\label{thm:sharp-kernel-unitarity-optical-theorem}
Let \(S=\mathbf 1+\ii T\) be unitary on the asymptotic Hilbert space.  With
the amplitude convention of Definition~\ref{def:invariant-amplitude-convention},
the connected sharp kernels obey
\[
  -\ii\left[
  \mathcal M(f|i)-\mathcal M(i|f)^*
  \right]
  =
  \sum_X
  \int\dd\Phi_X\,
  \mathcal M(X|f)^*\mathcal M(X|i),
\]
where \(X\) runs over a complete orthogonal resolution of outgoing
asymptotic channels, including the corresponding phase-space measures and
identical-particle symmetry factors.  In the forward case \(f=i\),
\[
  2\,\operatorname{Im}\mathcal M(i|i)
  =
  \sum_X
  \int\dd\Phi_X\,|\mathcal M(X|i)|^2 .
\]
For two incoming particles,
\[
  \sigma_{\rm tot}
  =
  \frac{2\,\operatorname{Im}\mathcal M(i|i)}
  {\mathcal F(p_1,p_2)} .
\]
\end{theorem}

\begin{proof}
The operator identity
\[
  S^\dagger S=\mathbf 1,\qquad S=\mathbf 1+\ii T,
\]
is equivalent to
\[
  -\ii(T-T^\dagger)=T^\dagger T .
\]
Let \(i\) and \(f\) denote incoming and outgoing asymptotic momentum labels,
including all discrete quantum numbers.  Insert a complete set of outgoing
asymptotic states between \(T^\dagger\) and \(T\).  Removing the common
overall momentum-conservation delta function gives the first displayed
sharp-kernel identity.
The conjugated factor on the right is fixed by Hilbert-space adjunction:
\[
  \langle f|T^\dagger|X\rangle
  =
  \langle X|T|f\rangle^* .
\]
Thus it is \(\mathcal M(X|f)^*\), not in general
\(\mathcal M(f|X)^*\).  Replacing one by the other would require an additional
reciprocity statement, for example from a time-reversal or detailed-balance
symmetry after its antiunitary phases have been fixed.

Setting \(f=i\) gives the optical theorem.
No time-reversal invariance is used in this forward specialization:
\(\langle i|T^\dagger|X\rangle=\langle X|T|i\rangle^*\) by the definition of
the adjoint, so each term is the norm square of the outgoing transition
amplitude from the fixed incoming state.  Dividing the complete transition
probability per unit time by the incoming flux gives the total cross section.
The amplitude on the left is the forward amplitude: the outgoing labels are
the same as the incoming labels.
\end{proof}

\begin{remark}[Complete sums and detector resolution]
The symbol \(X\) in the optical theorem denotes a complete orthogonal
resolution of the outgoing scattering Hilbert space in the ideal theory under
discussion.  A detector with finite resolution is represented more accurately
by a positive outgoing effect \(E_{\mathcal C}\), \(0\le E_{\mathcal C}\le
\mathbf 1\), associated with an observed event class \(\mathcal C\).  The
corresponding measured transition probability is
\[
  \langle i|T^\dagger E_{\mathcal C}T|i\rangle ,
\]
or, in sharp-kernel notation, the same phase-space sum as above with the
weights and integration region specified by \(E_{\mathcal C}\).  The forward
imaginary part \(2\,\operatorname{Im}\mathcal M(i|i)\) corresponds to the
special choice \(E_{\mathcal C}=\mathbf 1\), namely the complete inclusive sum
over all exact outgoing states of the theory.  Angular cuts, energy cuts, jet
definitions, and unresolved soft radiation define different effects
\(E_{\mathcal C}\) and hence different cross sections.

In a theory with massless long-range fields the complete sharp
fixed-particle-number terms are infrared-regulator dependent.  The operational
quantity is obtained by first keeping an infrared regulator, summing over the
states that the detector cannot distinguish, and only then removing the
regulator.  Chapter~\ref{ch:volume-ii-infrared-inclusive-qed} carries this out
for charged QED, where real unresolved photons and virtual soft photons combine
into a finite resolution-inclusive probability.
\end{remark}

\section{Partial-Wave States in Four Dimensions}

Partial waves are most transparently defined as generalized eigenstates of the
free two-particle asymptotic representation.  In the center-of-mass frame use
the \(\delta\)-normalized two-particle basis and write, for scalar particles,
\[
  {}_\delta\langle \vec k_1,\vec k_2
  |E,\vec P=\vec0,\ell,m\rangle
  =
  \delta^{(3)}(\vec k_1+\vec k_2)\,
  \delta(\omega_1+\omega_2-E)\,
  \mathcal N(E)Y_{\ell m}(\widehat k_1).
\]
The angular momentum labels satisfy
\[
  J^2=\ell(\ell+1),\qquad J_3=m.
\]
For identical spinless bosons, only even \(\ell\) occurs.  The normalization is
chosen so that
\[
  \langle E',\vec P',\ell',m'|E,\vec P,\ell,m\rangle
  =
  \delta(E'-E)\delta^{(3)}(\vec P'-\vec P)
  \delta_{\ell'\ell}\delta_{m'm}.
\]
Using
\[
  \int k^2\dd k\,\delta(\omega_1(k)+\omega_2(k)-E)
  =
  \frac{k\,\omega_1\omega_2}{E},
\]
and using the \(1/2\) in the symmetric two-particle resolution of the identity,
one finds
\[
  \mathcal N(E)
  =
  \left(\frac{2E}{k\,\omega_1\omega_2}\right)^{1/2}
  \qquad\text{for identical scalar bosons}.
\]
For distinguishable particles the same formula lacks the factor of \(2\) under
the square root.

Rotational invariance in the center-of-mass frame makes the elastic partial
waves diagonal:
\[
  S(E',\vec P',\ell',m'|E,\vec P,\ell,m)
  =
  \delta(E'-E)\delta^{(3)}(\vec P'-\vec P)
  \delta_{\ell'\ell}\delta_{m'm}\,S_\ell(E)
\]
in an elastic single-channel region.  If inelastic channels are present, this
equation describes the elastic matrix element of a larger unitary matrix, so
\[
  |S_\ell(E)|\le 1,
\]
with equality in a purely elastic partial wave.

In the same \(\delta\)-normalized convention, the connected elastic
sharp-momentum kernel is obtained by inserting the resolution of the identity
in partial waves.  For identical scalar bosons in the center-of-mass frame,
\[
  \ii(2\pi)^4M_\delta(\widehat q|\widehat p)
  =
  \frac{\mathcal N(E)^2}{4\pi}
  \sum_{\substack{\ell=0\\ \ell\ {\rm even}}}^{\infty}
  (2\ell+1)P_\ell(\widehat q\cdot\widehat p)
  \bigl(S_\ell(E)-1\bigr).
\]
This formula is the direct bridge between the partial-wave diagonalization of
the \(S\)-operator and the sharp-momentum amplitude.  Multiplying by the
external normalization factors converts \(M_\delta\) into the invariant
amplitude \(\mathcal M\) used in the cross-section formula.

\section{Partial-Wave Amplitudes in Four Dimensions}

For scalar \(2\to2\) scattering in the center-of-mass frame, rotational
invariance makes the amplitude a function of \(s\) and
\[
  x=\cos\theta=\widehat p_*\cdot\widehat q_* .
\]
For distinguishable particles, or for an ordered pair of external labels, use
the partial-wave expansion
\[
  \mathcal M(s,x)
  =
  16\pi
  \sum_{\ell=0}^{\infty}
  (2\ell+1)a_\ell(s)P_\ell(x),
\]
where \(P_\ell\) is the Legendre polynomial.  The inverse formula is
\[
  a_\ell(s)
  =
  \frac{1}{32\pi}
  \int_{-1}^1\dd x\,
  P_\ell(x)\mathcal M(s,x).
\]
This section fixes the ordered \(16\pi\) normalization.  For an unordered
identical-boson channel the physical amplitude is the Bose symmetrization of
the ordered amplitude; the corresponding even-\(\ell\) \(32\pi\) convention is
displayed below.
For equal incoming masses in the elastic region, define
\[
  \beta(s)=\frac{2|\vec p_*|}{\sqrt s}.
\]
This \(\beta(s)\) is a scalar two-body kinematic factor.  The letter
\(\rho\) is reserved in this monograph for spectral measures and their
densities, such as the K\"all\'en--Lehmann measure \(d\rho\).
The partial-wave \(S\)-matrix eigenvalue is
\[
  S_\ell(s)=1+2\ii\beta(s)a_\ell(s).
\]
Elastic unitarity is
\[
  |S_\ell(s)|=1.
\]
Equivalently, there is a real phase shift \(\delta_\ell(s)\) such that
\[
  S_\ell(s)=\ee^{2\ii\delta_\ell(s)}
\]
and
\[
  a_\ell(s)
  =
  \frac{\ee^{2\ii\delta_\ell(s)}-1}{2\ii\beta(s)}
  =
  \frac{\ee^{\ii\delta_\ell(s)}\sin\delta_\ell(s)}{\beta(s)} .
\]
The same condition can be written
\[
  \operatorname{Im}a_\ell(s)=\beta(s)|a_\ell(s)|^2 .
\]
With inelastic channels open,
\[
  |S_\ell(s)|\le1,
\]
with equality precisely in a purely elastic partial wave.

For distinguishable scalar particles, the elastic total cross section is
\[
  \sigma_{\rm el}
  =
  \frac{\pi}{|\vec p_*|^2}
  \sum_{\ell=0}^\infty
  (2\ell+1)|S_\ell(s)-1|^2
  =
  \frac{4\pi}{|\vec p_*|^2}
  \sum_{\ell=0}^\infty
  (2\ell+1)\sin^2\delta_\ell(s).
\]

For identical scalar bosons, the physical two-particle amplitude on unordered
states is Bose symmetric.  With the convention
\[
  \mathcal M_{\rm id}(s,x)
  =
  32\pi
  \sum_{\substack{\ell=0\\ \ell\ {\rm even}}}^{\infty}
  (2\ell+1)a_\ell(s)P_\ell(x),
\]
and the full-solid-angle final-state factor \(1/2\), the elastic cross
section is
\[
  \sigma_{\rm el}^{\rm id}
  =
  \frac{2\pi}{|\vec p_*|^2}
  \sum_{\substack{\ell=0\\ \ell\ {\rm even}}}^{\infty}
  (2\ell+1)|S_\ell(s)-1|^2 .
\]
This is the convention used later for identical-particle high-energy bounds.

\section{Partial-Wave Unitarity Geometry}

\paragraph{Partial-wave unitarity circle.}
Define
\[
  b_\ell(s)=\beta(s)a_\ell(s),
  \qquad
  S_\ell(s)=1+2\ii b_\ell(s).
\]
In an elastic partial wave, \(|S_\ell|=1\).  Writing
\[
  b_\ell=x_\ell+\ii y_\ell,
\]
this condition becomes
\[
  x_\ell^2+\left(y_\ell-\frac12\right)^2=\frac14 .
\]
Thus the elastic amplitude lies on the circle of radius \(1/2\) centered at
\(\ii/2\) in the complex \(b_\ell\)-plane.  Equivalently,
\[
  y_\ell=|b_\ell|^2 .
\]
The phase-shift form
\[
  b_\ell=\ee^{\ii\delta_\ell}\sin\delta_\ell
\]
parametrizes this circle.

When inelastic channels are open, the elastic partial wave is a matrix
element of a unitary matrix on a larger channel space.  It is convenient to
write
\[
  S_\ell=\eta_\ell\ee^{2\ii\delta_\ell},
  \qquad
  0\le\eta_\ell\le1 .
\]
Then \(b_\ell=(S_\ell-1)/(2\ii)\) lies inside the same circle:
\[
  x_\ell^2+\left(y_\ell-\frac12\right)^2\le\frac14,
  \qquad
  y_\ell\ge |b_\ell|^2 .
\]
The difference
\[
  y_\ell-|b_\ell|^2
  =
  \frac{1-\eta_\ell^2}{4}
\]
is the contribution of inelastic final states in that angular-momentum
channel.  This is the partial-wave form of the optical theorem.

Substitute \(S_\ell=1+2\ii b_\ell\) with
\(b_\ell=x_\ell+\ii y_\ell\).  Then
\[
  S_\ell=(1-2y_\ell)+2\ii x_\ell .
\]
The elastic condition \(|S_\ell|=1\) is therefore
\[
  (1-2y_\ell)^2+4x_\ell^2=1,
\]
which is equivalent to the displayed circle equation and to
\(y_\ell=|b_\ell|^2\).  If \(S_\ell=\eta_\ell e^{2\ii\delta_\ell}\) with
\(0\le\eta_\ell\le1\), the same substitution gives
\[
  \left|1+2\ii b_\ell\right|^2=\eta_\ell^2 .
\]
Rearranging yields
\[
  y_\ell-|b_\ell|^2=\frac{1-\eta_\ell^2}{4}\ge0,
\]
and the inequality form of the unitarity disk follows.  The right side is the
deficit \(1-\eta_\ell^2\) in the norm of the elastic component, expressed in
the \(b_\ell=(S_\ell-1)/(2\ii)\) normalization.  This is the partial-wave
optical theorem.

\section{Thresholds, Poles, and Resonances}

The same variables organize the analytic structure of scattering.  A
threshold is the value of \(s\) at which a new on-shell intermediate channel
becomes kinematically available.  Across such a threshold, unitarity gives
the discontinuity of amplitudes in terms of phase-space integrals over lower
amplitudes.

A stable bound state in a two-particle channel appears as a pole of the
partial-wave amplitude on the physical, or first, sheet below threshold.  A
resonance is represented by a pole reached after analytic continuation through
a unitarity cut to an adjacent sheet.  The sheet terminology and the distinction
between a physical real region and the physical analytic sheet are fixed in
Section~\ref{sec:physical-region-and-first-sheet}; the organizing data are
already present here: the \(S\)-operator, phase space, unitarity, and partial
waves.
