\chapter{Massless Particles, Helicity, and Gauge Redundancy}
\label{ch:massless-particles-helicity-gauge-redundancy}
\ConventionsLorentzian


\section{The Massless Orbit}

Work in \(D=4\) with mostly-plus metric
\[
  \eta_{\mu\nu}=\operatorname{diag}(-,+,+,+).
\]
The positive-energy massless orbit is
\[
  \Sigma_0^+
  =
  \{k=(k^0,\vec k):k^2=0,\ k^0=|\vec k|,\ k^0>0\}.
\]
Together with the massive positive-energy shells of
Chapter~\ref{ch:massive-particles-spin}, this is one of the orbit types
compatible with the spectrum condition; the complete four-dimensional orbit
classification is summarized in
Table~\ref{tab:wigner-classification-four-dimensional}.  Spacelike Wigner
orbits \(k^2>0\) are excluded by
\(\operatorname{Spec}(P)\subset\overline V_+\), as stated there.
Choose a reference momentum
\[
  q=(\kappa,0,0,\kappa),
  \qquad
  \kappa>0.
\]
For each \(k\in\Sigma_0^+\), choose a standard Lorentz transformation
\[
  L(k)\in SO^+(1,3),
  \qquad
  L(k)q=k.
\]
As in the massive case, Hilbert-space formulas use the universal cover of
\(SO^+(1,3)\) when spinorial phases are involved, while the same symbols
denote the projected Lorentz transformations acting on four-vectors.
The little group of \(q\) is
\[
  G_q=\{W\in SO^+(1,3):Wq=q\}.
\]

\begin{definition}[Massless one-particle orbit and standard section]
\label{def:massless-orbit-standard-section}
  The positive-energy massless orbit is the homogeneous space
  \(\Sigma_0^+\simeq SO^+(1,3)/G_q\), where \(G_q\) is the stabilizer of
  \(q=(\kappa,0,0,\kappa)\).  A standard section is a choice of Lorentz
  transformation \(L(k)\) for each \(k\in\Sigma_0^+\) such that \(L(k)q=k\).
  It is not canonical: replacing \(L(k)\) by \(L(k)Q(k)\), with
  \(Q(k)\in G_q\), changes the one-particle helicity frame.
\end{definition}

The non-uniqueness of \(L(k)\) is the origin of helicity-frame phases and of
the representative freedom of vector polarizations.  It is the choice of
local trivialization of a line bundle over the null orbit.

For vector indices, write a Lorentz transformation as
\[
  \Lambda
  =
  \exp\!\left(\frac{\ii}{2}\omega_{\mu\nu}M^{\mu\nu}\right),
  \qquad
  \omega_{\mu\nu}=-\omega_{\nu\mu},
\]
where
\[
  (M^{\mu\nu})^\rho{}_\sigma
  =
  -\ii\left(
    \eta^{\mu\rho}\delta^\nu_\sigma
    -
    \eta^{\nu\rho}\delta^\mu_\sigma
  \right).
\]
Then, for infinitesimal \(\omega\),
\[
  \Lambda^\mu{}_\nu=\delta^\mu{}_\nu+\omega^\mu{}_\nu,
\]
and
\[
  [M^{\mu\nu},M^{\rho\sigma}]
  =
  -\ii\left(
    \eta^{\mu\rho}M^{\nu\sigma}
    -
    \eta^{\nu\rho}M^{\mu\sigma}
    -(\rho\leftrightarrow\sigma)
  \right).
\]
Let \(J_{\mu\nu}\) denote the corresponding Hilbert-space generators on the
one-particle representation:
\[
  U(\Lambda)
  =
  \exp\!\left(\frac{\ii}{2}\omega_{\mu\nu}J^{\mu\nu}\right),
  \qquad
  [J^{\mu\nu},J^{\rho\sigma}]
  =
  -\ii\left(
    \eta^{\mu\rho}J^{\nu\sigma}
    -
    \eta^{\nu\rho}J^{\mu\sigma}
    -(\rho\leftrightarrow\sigma)
  \right).
\]
The three Hilbert-space generators
\[
  J_3=J_{12},
  \qquad
  N_1=J_{01}+J_{31},
  \qquad
  N_2=J_{02}+J_{32}
\]
generate \(G_q\). Their commutators are
\[
  [N_1,N_2]=0,
  \qquad
  [J_3,N_1]=\ii N_2,
  \qquad
  [J_3,N_2]=-\ii N_1.
\]

\begin{proposition}[Little algebra of a future null momentum]
\label{prop:massless-little-algebra-iso2}
  The infinitesimal stabilizer of \(q=(\kappa,0,0,\kappa)\) is generated by
  \(J_3=J_{12}\), \(N_1=J_{01}+J_{31}\), and \(N_2=J_{02}+J_{32}\), and these
  generators obey
  \[
    [N_1,N_2]=0,\qquad
    [J_3,N_1]=\ii N_2,\qquad
    [J_3,N_2]=-\ii N_1 .
  \]
  Hence the connected little group is the double cover of
  \(\mathbb R^2\rtimes SO(2)\).
\end{proposition}

\begin{proof}
In the four-vector representation the matrices \(M_{12}\),
\(M_{01}+M_{31}\), and \(M_{02}+M_{32}\) annihilate \(q\).  Conversely, if
an infinitesimal Lorentz transformation \(\omega^\mu{}_\nu\) stabilizes
\(q\), then \(\omega q=0\).  With \(q^0=q^3=\kappa\), this condition and the
\(\eta\)-antisymmetry of \(\omega\) leave precisely one rotation parameter in
the \(1,2\)-plane and two null-translation parameters, represented by the
three generators displayed above.

The commutators follow by substituting these linear combinations into the
Lorentz algebra.  The translation commutator cancels between the
\((01,02)\) and \((31,32)\) contributions, while \(J_{12}\) rotates the pair
\((J_{01}+J_{31},J_{02}+J_{32})\) as a two-vector.  This is the Lie algebra
of the Euclidean group of the transverse plane.
\end{proof}

This is the Lie algebra of the Euclidean group \(ISO(2)\): \(J_3\) generates
rotations in a two-dimensional transverse plane, while \(N_1,N_2\) generate
commuting translations in that plane. The double cover is used on Hilbert
space.

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.95]
  \draw[->,line width=0.7pt] (-3.0,0)--(3.2,0)
    node[right] {translation subgroup};
  \draw[->,line width=0.7pt] (0,-1.9)--(0,2.15)
    node[above] {rotation};
  \node[blue!70!black] at (0,1.35) {$J_3$};
  \draw[blue!70!black,very thick,->]
    (0.55,0.25) arc[start angle=25,end angle=325,radius=0.58];
  \node[green!50!black] at (-2.05,-0.35) {$N_1$};
  \node[green!50!black] at (2.05,-0.35) {$N_2$};
  \draw[green!50!black,very thick,->] (-1.55,-0.85)--(-0.45,-0.85);
  \draw[green!50!black,very thick,->] (0.45,-0.85)--(1.55,-0.85);
  \node[align=center] at (0,-1.45)
    {$ISO(2)=\mathbb R^2\rtimes SO(2)$};
\end{tikzpicture}
\end{center}

\section{Little-Group Matrices}

The matrix representatives make the \(ISO(2)\) structure explicit. The
rotation around the \(x^3\)-axis is generated by
\[
  (M_{12})^\mu{}_\nu
  =
  \begin{pmatrix}
    0&0&0&0\\
    0&0&-\ii&0\\
    0&\ii&0&0\\
    0&0&0&0
  \end{pmatrix},
\]
so
\[
  R_3(\theta)
  =
  \ee^{\ii\theta M_{12}}
  =
  \begin{pmatrix}
    1&0&0&0\\
    0&\cos\theta&\sin\theta&0\\
    0&-\sin\theta&\cos\theta&0\\
    0&0&0&1
  \end{pmatrix}.
\]
The two null-translation generators are
\[
  A=M_{01}+M_{31}
  =
  -\ii
  \begin{pmatrix}
    0&1&0&0\\
    1&0&0&-1\\
    0&0&0&0\\
    0&1&0&0
  \end{pmatrix},
\]
and
\[
  B=M_{02}+M_{32}
  =
  -\ii
  \begin{pmatrix}
    0&0&1&0\\
    0&0&0&0\\
    1&0&0&-1\\
    0&0&1&0
  \end{pmatrix}.
\]
Since \(A^3=B^3=A^2B=AB^2=0\) in this representation and
\([A,B]=0\), their exponential is the finite polynomial
\[
  T(\alpha,\beta)
  =
  \ee^{\ii\alpha A+\ii\beta B}
  =
  \begin{pmatrix}
    1+\zeta&\alpha&\beta&-\zeta\\
    \alpha&1&0&-\alpha\\
    \beta&0&1&-\beta\\
    \zeta&\alpha&\beta&1-\zeta
  \end{pmatrix},
  \qquad
  \zeta=\frac{\alpha^2+\beta^2}{2}.
\]
Every displayed matrix fixes \(q\), and the little-group element with
parameters \((\alpha,\beta,\theta)\) is
\[
  W(\alpha,\beta,\theta)
  =
  T(\alpha,\beta)R_3(\theta).
\]

\begin{proposition}[Null-translation matrices]
\label{prop:null-translation-matrices}
  The matrix \(T(\alpha,\beta)\) satisfies
  \[
    T(\alpha,\beta)^T\eta\,T(\alpha,\beta)=\eta,\qquad
    T(\alpha,\beta)q=q ,
  \]
  and the product law
  \[
    T(\alpha,\beta)T(\alpha',\beta')
    =
    T(\alpha+\alpha',\beta+\beta').
  \]
  The rotation \(R_3(\theta)\) acts on the pair
  \((\alpha,\beta)\) by the ordinary rotation of the transverse plane.
\end{proposition}

\begin{proof}
The displayed \(T(\alpha,\beta)\) is the exponential of
\(\ii\alpha A+\ii\beta B\).  Since \(A\) and \(B\) are Lorentz Lie-algebra
matrices, the exponential preserves \(\eta\).  Multiplication of the displayed
matrix by \((\kappa,0,0,\kappa)^T\) gives \((\kappa,0,0,\kappa)^T\), so
\(T(\alpha,\beta)\) fixes \(q\).  Because \(A\) and \(B\) commute, their
exponentials multiply by adding parameters.  The commutators
\([M_{12},A]=\ii B\) and \([M_{12},B]=-\ii A\) give the stated rotational
action on the null-translation parameters.
\end{proof}

The vector-space commutators
\[
  [M_{12},A]=\ii B,
  \qquad
  [M_{12},B]=-\ii A
\]
are the matrix form of the semidirect product
\(\mathbb R^2\rtimes SO(2)\).

\section{Helicity Representations}

The Hilbert-space counterparts of \(A\) and \(B\) are \(N_1\) and \(N_2\).
Since \([N_1,N_2]=0\), a fixed-\(q\) generalized basis, in the sense of
Section~\ref{sec:rigged-hilbert-continuous-spectrum}, may be chosen so that
\[
  N_1\ket{q,a,b;\lambda}=a\ket{q,a,b;\lambda},
  \qquad
  N_2\ket{q,a,b;\lambda}=b\ket{q,a,b;\lambda}.
\]
Here \(\lambda\) denotes any remaining degeneracy label. The commutators with
\(J_3\) imply that, on these labels,
\[
  J_3
  =
  \ii b\frac{\partial}{\partial a}
  -
  \ii a\frac{\partial}{\partial b}
  +
  J_3^{(0)},
\]
where \(J_3^{(0)}\) acts on the degeneracy space at fixed \((a,b)\).
Consequently
\[
  \ee^{\ii\theta J_3}\ket{q,a,b;\lambda}
  =
  \ket{q,a\cos\theta-b\sin\theta,
          a\sin\theta+b\cos\theta;\lambda_\theta}
\]
up to the action of \(J_3^{(0)}\). If \(a^2+b^2>0\), the one-particle
labels at fixed momentum include the full \(SO(2)\)-orbit of \((a,b)\).
Those irreducible representations are continuous-spin representations. The
helicity sectors are the massless one-particle irreducible representations
for which
\[
  N_1\ket{q,h}=0,
  \qquad
  N_2\ket{q,h}=0.
\]
The remaining little-group generator acts by
\[
  J_3\ket{q,h}=h\ket{q,h}.
\]
Equivalently, for
\[
  W(\alpha,\beta,\theta)
  =
  \exp(\ii\alpha N_1+\ii\beta N_2)\exp(\ii\theta J_3),
\]
the state transforms as
\[
  U(W(\alpha,\beta,\theta))\ket{q,h}
  =
  \ee^{\ii h\theta}\ket{q,h}.
\]
The double-cover rotation condition gives
\[
  h\in\frac12\mathbb Z.
\]

\begin{definition}[Massless helicity representation]
\label{def:massless-helicity-representation}
  A massless one-particle helicity representation is the induced
  positive-energy Poincare representation built from the little-group
  character
  \[
    \chi_h(W(\alpha,\beta,\theta))=\ee^{\ii h\theta},
    \qquad h\in\frac12\mathbb Z ,
  \]
  in which the translation subgroup \(\mathbb R^2\subset ISO(2)\) acts
  trivially.  Equivalently, at the reference momentum \(q\),
  \(N_1=N_2=0\) and \(J_3\) acts by the scalar \(h\).
\end{definition}

\begin{proposition}[Continuous-spin alternative]
\label{prop:continuous-spin-little-group-orbit}
  If the joint spectrum of the commuting little-group translations contains
  a point \((a,b)\) with \(a^2+b^2=\rho^2>0\), irreducibility under the
  little group forces the whole circle of radius \(\rho\) in the
  \((a,b)\)-plane to occur.  The corresponding massless representation has
  infinitely many \(SO(2)\) Fourier modes at fixed momentum and is not a
  finite-helicity representation.
\end{proposition}

\begin{proof}
The commutators of Proposition~\ref{prop:massless-little-algebra-iso2} imply
that \(J_3\) generates rotations of the joint translation spectrum.  Therefore
the spectral point \((a,b)\) is accompanied by
\((a\cos\theta-b\sin\theta,a\sin\theta+b\cos\theta)\) for every
\(\theta\).  If \(\rho>0\), the orbit is a circle.  The Hilbert space on this
circle carries the regular \(SO(2)\) action, whose Fourier decomposition has
integer-spaced modes.  Hence the little-group representation contains an
infinite tower of angular modes rather than a single helicity character.
\end{proof}

\begin{remark}[Continuous-spin sectors]
The alternative \(a^2+b^2=\rho^2>0\) gives the massless continuous-spin, or
infinite-spin, Wigner representations.  They are legitimate irreducible
positive-energy representations of the Poincare group, so the restriction to
\(\rho=0\) is an additional physical and structural choice.  At fixed momentum
the little-group orbit is a circle, and the Fourier modes on that circle form
an infinite tower of helicity labels mixed by boosts.  A gas of such particles
therefore has infinitely many internal polarizations at each momentum; ordinary
finite-helicity thermodynamic counting is lost unless further structure
controls the occupation of those modes.

There is also a sharper locality obstruction.  Under the usual assumptions of
positive energy, covariance, local commutativity, and pointlike covariant local
fields, the zero-mass infinite-spin representations cannot be interpolated by a
local field with a nonzero vacuum-to-one-particle matrix element.  Equivalently,
their natural localization in algebraic constructions is spacelike-cone or
string localization rather than point localization in bounded regions.  Such
sectors are therefore possible representation-theoretic options, but they do
not enter the point-local or gauge-redundant helicity field constructions used
for photons, gravitons, and the gauge theories developed in this volume.
\label{rem:continuous-spin-scope}
\end{remark}

There are two useful normalization conventions. The covariant convention used
below is
\[
  \langle k',h'|k,h\rangle
  =
  (2\pi)^3\,2k^0\,\delta^{(3)}(\vec k'-\vec k)\delta_{h'h}.
\]
In the delta-normalized convention
\[
  {}_\delta\!\langle \vec k',h'|\vec k,h\rangle_\delta
  =
  \delta^{(3)}(\vec k'-\vec k)\delta_{h'h},
\]
the corresponding state is
\[
  \ket{\vec k,h}_\delta
  =
  \sqrt{\frac{q^0}{k^0}}\,
  U(L(k))\ket{\vec q,h}_\delta .
\]
The factor is fixed by the Lorentz transformation of the three-dimensional
delta function on the mass shell. It gives
\[
  U(\Lambda)\ket{\vec k,h}_\delta
  =
  \sqrt{\frac{(\Lambda k)^0}{k^0}}\,
  \ee^{\ii h\theta(\Lambda,k)}
  \ket{\overrightarrow{\Lambda k},h}_\delta .
\]

For general \(k\in\Sigma_0^+\), define generalized one-particle states by
\[
  \ket{k,h}=U(L(k))\ket{q,h},
\]
with the covariant normalization displayed above.
For a Lorentz transformation \(\Lambda\), the Wigner little-group element is
\[
  W(\Lambda,k)=L(\Lambda k)^{-1}\Lambda L(k),
\]
and the helicity state transforms by
\[
  U(\Lambda)\ket{k,h}
  =
  \ee^{\ii h\theta(\Lambda,k)}\ket{\Lambda k,h},
\]
where \(\theta(\Lambda,k)\) is the \(SO(2)\) rotation angle of
\(W(\Lambda,k)\) in the helicity representation. Different choices of
\(L(k)\) change the phase convention for \(\ket{k,h}\) but leave invariant
matrix elements unchanged.

\section{Helicity Frames}

For a fixed helicity \(h\), let \(V_h\simeq \mathbb C\) be the one-dimensional
little-group representation with character
\[
  \chi_h(W(\alpha,\beta,\theta))=\ee^{\ii h\theta}.
\]
The translation parameters \(\alpha,\beta\) do not appear in \(\chi_h\)
because this is the helicity sector. With
\[
  d\mu_0(k)=\frac{d^3k}{(2\pi)^3\,2|\vec k|},
\]
the one-particle Hilbert space of a single helicity is represented, after
choosing \(L(k)\), as
\[
  \Hilb_{1,h}\simeq L^2(\Sigma_0^+,d\mu_0;V_h).
\]
The distributional frame \(F_L(k):V_h\to\Hilb_{1,h}\) satisfies
\[
  U(\Lambda)F_L(k)
  =
  F_L(\Lambda k)\chi_h(W(\Lambda,k)).
\]
The little-group element again obeys
\[
  W(\Lambda_2\Lambda_1,k)
  =
  W(\Lambda_2,\Lambda_1 k)W(\Lambda_1,k),
\]
so the helicity phase is a one-dimensional cocycle.

\begin{proposition}[Wigner cocycle and change of helicity frame]
\label{prop:wigner-cocycle-helicity-frame-change}
  For a standard section \(L(k)\), the little-group element
  \(W(\Lambda,k)=L(\Lambda k)^{-1}\Lambda L(k)\) obeys
  \[
    W(\Lambda_2\Lambda_1,k)
    =
    W(\Lambda_2,\Lambda_1 k)W(\Lambda_1,k).
  \]
  If \(L'(k)=L(k)Q(k)\), then
  \[
    W'(\Lambda,k)=Q(\Lambda k)^{-1}W(\Lambda,k)Q(k),
  \]
  and the corresponding helicity-frame wavefunctions obey
  \[
    \psi_{L'}(k)=\chi_h(Q(k))^{-1}\psi_L(k).
  \]
\end{proposition}

\begin{proof}
The cocycle identity is the associative identity
\[
  L(\Lambda_2\Lambda_1k)^{-1}\Lambda_2\Lambda_1L(k)
  =
  \bigl(L(\Lambda_2\Lambda_1k)^{-1}\Lambda_2L(\Lambda_1k)\bigr)
  \bigl(L(\Lambda_1k)^{-1}\Lambda_1L(k)\bigr).
\]
For \(L'(k)=L(k)Q(k)\), substitution gives
\[
  L'(\Lambda k)^{-1}\Lambda L'(k)
  =
  Q(\Lambda k)^{-1}L(\Lambda k)^{-1}\Lambda L(k)Q(k),
\]
which is the displayed conjugation law.  The distributional frame changes by
right action of the one-dimensional representation,
\[
  F_{L'}(k)=F_L(k)\chi_h(Q(k)).
\]
Since a vector is written as
\[
  \int_{\Sigma_0^+} F_L(k)\psi_L(k)\,d\mu_0(k),
\]
invariance of this vector under the frame change forces the inverse character
on the wavefunction component.
\end{proof}

\section{Photon Intertwiners and Vector Representatives}

Consider helicity \(h=\pm1\). A vector polarization representative at the
reference momentum is
\[
  \epsilon_\mu^+(q)=\frac1{\sqrt2}(0,1,-\ii,0),
  \qquad
  \epsilon_\mu^-(q)=\frac1{\sqrt2}(0,1,\ii,0).
\]
It obeys
\[
  q^\mu\epsilon_\mu^\pm(q)=0,
  \qquad
  \epsilon_\mu^\pm(q)\epsilon^{\pm *\,\mu}(q)=1,
  \qquad
  \epsilon_\mu^\pm(q)\epsilon^{\mp *\,\mu}(q)=0.
\]
The rotation generated by \(J_3\) gives
\[
  W(0,0,\theta)^\nu{}_\mu\epsilon_\nu^\pm(q)
  =
  \ee^{\pm\ii\theta}\epsilon_\mu^\pm(q),
\]
which is the vector-representative phase entering the covariance equation;
the one-particle state contributes the inverse little-group phase.
The sign in this equation follows from the lower-index action.  With the
rotation matrix displayed above, restricted to the transverse \(1,2\)
components,
\[
\begin{aligned}
  R_3(\theta)^\nu{}_\mu\epsilon_\nu^+(q)
  &=
  \frac1{\sqrt2}
  \bigl(0,\cos\theta+\ii\sin\theta,\,
          \sin\theta-\ii\cos\theta,0\bigr)  \\
  &=
  \ee^{\ii\theta}\epsilon_\mu^+(q),
\end{aligned}
\]
and the complex conjugate vector \(\epsilon_\mu^-(q)\) gives the phase
\(\ee^{-\ii\theta}\).

Let \(\widehat\Phi_\mu\) be a covariant vector local field, in a
positive-metric physical Hilbert space, with a nonzero one-photon matrix
element. In the delta-normalized convention write
\[
  {}_\delta\!\langle \vec k,h|\widehat\Phi_\mu(0)\ket{\vac}
  =
  C_h u_\mu^h(\vec k).
\]
If
\[
  U(\Lambda)\widehat\Phi_\mu(x)U(\Lambda)^{-1}
  =
  \Lambda^\nu{}_\mu\,\widehat\Phi_\nu(\Lambda x),
\]
then the transformation law of the delta-normalized states gives
\[
  u_\mu^h(\vec k)
  =
  \sqrt{\frac{(\Lambda k)^0}{k^0}}\,
  \ee^{-\ii h\theta(\Lambda,k)}
  \Lambda^\nu{}_\mu u_\nu^h(\Lambda\vec k).
\]
At \(k=q\) and \(\Lambda=W(\alpha,\beta,\theta)\), this specializes to
\[
  u_\mu^h(\vec q)
  =
  \ee^{-\ii h\theta}
  W^\nu{}_\mu u_\nu^h(\vec q).
\]
The \(SO(2)\) part of this equation fixes the transverse representatives
above. The null-translation part acts by a shift along \(q_\mu\).
For
\[
  W(\alpha,\beta,0)
  =
  \begin{pmatrix}
    1+\zeta&\alpha&\beta&-\zeta\\
    \alpha&1&0&-\alpha\\
    \beta&0&1&-\beta\\
    \zeta&\alpha&\beta&1-\zeta
  \end{pmatrix},
  \qquad
  \zeta=\frac{\alpha^2+\beta^2}{2},
\]
one finds
\[
  W(\alpha,\beta,0)^\nu{}_\mu\epsilon_\nu^\pm(q)
  =
  \frac{1}{\sqrt2}
  \begin{pmatrix}
    \alpha\mp\ii\beta\\
    1\\
    \mp\ii\\
    -\alpha\pm\ii\beta
  \end{pmatrix}
  =
  \epsilon_\mu^\pm(q)
  -
  \frac{\alpha\mp\ii\beta}{\sqrt2\,\kappa}\,q_\mu ,
\]
where \(q_\mu=(-\kappa,0,0,\kappa)\). Hence the vector representative is
defined modulo the null direction:
\[
  \epsilon_\mu^\pm(k)\sim \epsilon_\mu^\pm(k)+c\,k_\mu .
\]

\begin{proposition}[Little-group translations are polarization shifts]
\label{prop:little-group-translation-polarization-shift}
  For helicity \(h=\pm1\), the null-translation subgroup of the massless
  little group acts on the covariant vector representative by
  \[
    W(\alpha,\beta,0)^\nu{}_\mu\epsilon_\nu^\pm(q)
    =
    \epsilon_\mu^\pm(q)
    -
    \frac{\alpha\mp\ii\beta}{\sqrt2\,\kappa}\,q_\mu .
  \]
  Therefore the vector representative of a helicity-one state is naturally an
  equivalence class modulo \(k_\mu\)-shifts.
\end{proposition}

\begin{proof}
Multiplying \(T(\alpha,\beta)\) by \(\epsilon_\nu^\pm(q)\) gives the
four-component column displayed above.  Since
\(q_\mu=(-\kappa,0,0,\kappa)\), the difference between that column and
\(\epsilon_\mu^\pm(q)\) is
\(-(\alpha\mp\ii\beta)q_\mu/(\sqrt2\,\kappa)\).  For a general null momentum
\(k=L(k)q\), applying \(L(k)\) transports the same statement to a shift
proportional to \(k_\mu\).
\end{proof}

For general \(k\), choose
\[
  \epsilon_\mu^\pm(k)=L(k)_\mu{}^\nu\epsilon_\nu^\pm(q).
\]
If \(L'(k)=L(k)Q(k)\) and
\(Q(k)=W(\alpha(k),\beta(k),\theta(k))\), then
\[
  \epsilon_\mu^{\pm\,\prime}(k)
  =
  \ee^{\pm\ii\theta(k)}\epsilon_\mu^\pm(k)+c_\pm(k)k_\mu .
\]
The phase is the vector-representative helicity phase, and the term
proportional to \(k_\mu\) is the representative change generated by the
null-translation part of the little group.

\section{Field Strength Matrix Elements}

A vector potential \(A_\mu\) is a representative variable whose polarization
matrix element uses \(\epsilon_\mu^h(k)\). Its representative freedom is
\[
  A_\mu(x)\sim A_\mu(x)+\partial_\mu\xi(x),
\]
where \(\xi\) is a scalar gauge parameter. The gauge-invariant local tensor
constructed from \(A_\mu\) is
\[
  F_{\mu\nu}
  =
  \partial_\mu A_\nu-\partial_\nu A_\mu .
\]

For a one-photon state of helicity \(h=\pm1\), the field-strength matrix
element has the form
\[
  \langle k,h|\,\widehat F_{\mu\nu}(x)\,\ket{\vac}
  =
  C_h\ee^{-\ii k\cdot x}
  \left(
    -\ii k_\mu\epsilon_\nu^{h*}(k)
    +\ii k_\nu\epsilon_\mu^{h*}(k)
  \right),
\]
where \(C_h\) is a normalization constant fixed by the field normalization.
The shift
\[
  \epsilon_\mu^h(k)\mapsto\epsilon_\mu^h(k)+c\,k_\mu
\]
leaves this expression unchanged because
\[
  k_\mu k_\nu-k_\nu k_\mu=0.
\]
The local photon observable at this stage is therefore the field strength,
while the potential is a representative used to write local equations,
actions, and perturbative Green functions.

\begin{proposition}[Field-strength matrix element descends to the helicity line]
\label{prop:field-strength-descends-helicity-line}
  The one-photon field-strength matrix element
  \[
    \langle k,h|\,\widehat F_{\mu\nu}(x)\,\ket{\vac}
    =
    C_h\ee^{-\ii k\cdot x}
    \left(
      -\ii k_\mu\epsilon_\nu^{h*}(k)
      +\ii k_\nu\epsilon_\mu^{h*}(k)
    \right)
  \]
  is independent of the representative
  \(\epsilon_\mu^h(k)\) inside its equivalence class
  \(\epsilon_\mu^h(k)\sim\epsilon_\mu^h(k)+c\,k_\mu\).
\end{proposition}

\begin{proof}
Under the representative change, the extra contribution to the bracketed
tensor is proportional to
\[
  -\ii k_\mu k_\nu+\ii k_\nu k_\mu=0 .
\]
Thus the antisymmetric field strength is well-defined on the quotient of
vector representatives by the null direction.  Equivalently, the linear map
\[
  \epsilon_\mu\longmapsto k_\mu\epsilon_\nu-k_\nu\epsilon_\mu
\]
has the null line \(\mathbb C k_\mu\) in its kernel and therefore descends to
the quotient polarization line used for helicity-one states.  The
\(x\)-dependence and the normalization \(C_h\) are unaffected because the
representative change acts only on \(\epsilon_\mu^h(k)\).
\end{proof}

\begin{remark}
This statement is made in the physical positive-metric Hilbert space and its
local observable algebra. Gauge-fixed quantization later introduces
\(A_\mu\) on an enlarged space with constraints, ghosts, or BRST
cohomology. In that formulation the vector potential is a calculational
representative, and \(F_{\mu\nu}\) descends to the physical local algebra.
\end{remark}

\begin{remark}[Two external photons]
For an amplitude with an incoming photon of momentum \(k\) and an outgoing
photon of momentum \(k'\), write the reduced tensor before choosing
polarization representatives as \(\mathcal M^{\mu\nu}\).  Descent to the two
helicity lines requires two Ward identities,
\[
  k_\mu\mathcal M^{\mu\nu}=0,
  \qquad
  k'_\nu\mathcal M^{\mu\nu}=0 .
\]
In tree-level spinor Compton scattering these identities are obtained by
summing the two electron-exchange diagrams.  Let
\[
  N(q)=-\ii\not q+m,\qquad D(q)=q^2+m^2-\ii0 ,
\]
with \(\not q=\gamma^\rho q_\rho\),
and impose the external on-shell Dirac equations.  Replacing the incoming
polarization by \(k_\mu\) gives, up to the common charge and normalization
factor, the two terms
\[
  \bar u(p')\,\not\epsilon'{}^*(k')\,\frac{N(p+k)\not k}{D(p+k)}\,u(p)
  =
  -\ii\,\bar u(p')\not\epsilon'{}^*(k')u(p),
\]
and
\[
  \bar u(p')\,\not k\,\frac{N(p-k')}{D(p-k')}\,
  \not\epsilon'{}^*(k')\,u(p)
  =
  +\ii\,\bar u(p')\not\epsilon'{}^*(k')u(p).
\]
The contact terms produced by the Ward contraction of the \(s\)- and
\(u\)-channel diagrams therefore cancel.  Replacing the outgoing polarization
by \(k'_\nu\) gives the same cancellation with \(k\) and \(k'\) interchanged.
\end{remark}

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.95]
  \draw[->,gray!70,line width=0.7pt] (-3.4,0)--(3.6,0)
    node[right] {representatives};
  \draw[->,gray!70,line width=0.7pt] (0,-1.8)--(0,2.05)
    node[above] {$k_\mu$ shift};
  \fill[blue!70!black] (-1.35,0) circle (2.4pt)
    node[below left=0.12cm] {$\epsilon_\mu^h$};
  \fill[blue!70!black] (1.25,0.92) circle (2.4pt)
    node[above right=0.08cm] {$\epsilon_\mu^h+c k_\mu$};
  \draw[blue!70!black,very thick,->] (-1.35,0)--(1.25,0.92);
  \node[align=center,fill=white,inner sep=1.8pt] at (0,-1.28)
    {$k_\mu\epsilon_\nu^h-k_\nu\epsilon_\mu^h$\\is unchanged};
\end{tikzpicture}
\end{center}

\section{Polarization Sums}

For calculations with external helicity states it is useful to choose
representatives by imposing an auxiliary linear condition. Let \(n^\mu\) be a
fixed vector satisfying
\[
  n\cdot k\ne0 .
\]
Choose \(\epsilon_\mu^h(k;n)\) so that
\[
  k^\mu\epsilon_\mu^h(k;n)=0,
  \qquad
  n^\mu\epsilon_\mu^h(k;n)=0,
  \qquad
  \epsilon^h(k;n)\cdot\epsilon^{h'*}(k;n)=\delta_{hh'} .
\]
The sum over the two physical helicities is
\[
  \Pi_{\mu\nu}(k;n)
  =
  \sum_{h=\pm1}
  \epsilon_\mu^h(k;n)\epsilon_\nu^{h*}(k;n),
\]
with
\[
  \Pi_{\mu\nu}(k;n)
  =
  \eta_{\mu\nu}
  -
  \frac{k_\mu n_\nu+n_\mu k_\nu}{k\cdot n}
  +
  \frac{n^2 k_\mu k_\nu}{(k\cdot n)^2}.
\]
It satisfies
\[
  k^\mu\Pi_{\mu\nu}(k;n)=0,
  \qquad
  n^\mu\Pi_{\mu\nu}(k;n)=0.
\]
Changing \(n\) changes the representative polarization vectors by terms
proportional to \(k_\mu\). Therefore, if a quantity \(M^\mu(k)\) obeys
\[
  k_\mu M^\mu(k)=0,
\]
then
\[
  M^\mu(k)\Pi_{\mu\nu}(k;n)M^{\nu *}(k)
\]
is independent of the representative choice. This is the form in which
polarization sums will enter gauge-invariant scattering calculations.

\begin{proposition}[Auxiliary-vector polarization projector]
\label{prop:auxiliary-vector-polarization-projector}
  Let \(k^2=0\) and \(n\cdot k\ne0\).  The tensor
  \[
    \Pi_{\mu\nu}(k;n)
    =
    \eta_{\mu\nu}
    -
    \frac{k_\mu n_\nu+n_\mu k_\nu}{k\cdot n}
    +
    \frac{n^2 k_\mu k_\nu}{(k\cdot n)^2}
  \]
  annihilates \(k^\mu\) and \(n^\mu\).  If \(M^\mu k_\mu=0\), then the
  contraction \(M^\mu\Pi_{\mu\nu}(k;n)M^{\nu *}\) is unchanged under a change
  of polarization representatives by \(k_\mu\)-shifts.
\end{proposition}

\begin{proof}
Contracting with \(k^\mu\) gives
\[
  k_\nu-\frac{k^2 n_\nu+(k\cdot n)k_\nu}{k\cdot n}
  +\frac{n^2k^2k_\nu}{(k\cdot n)^2}=0,
\]
using \(k^2=0\).  Contracting with \(n^\mu\) gives
\[
  n_\nu-\frac{(n\cdot k)n_\nu+n^2k_\nu}{k\cdot n}
  +\frac{n^2(n\cdot k)k_\nu}{(k\cdot n)^2}=0.
\]
If the representative basis is changed, the projector changes only by terms
with at least one factor \(k_\mu\) or \(k_\nu\).  Such terms vanish in
\(M^\mu\Pi_{\mu\nu}M^{\nu *}\) when \(M^\mu k_\mu=0\).
\end{proof}

\section{Four-Dimensional Spinor-Helicity Variables}
\label{sec:spinor-helicity-variables}

The preceding sections described helicity as a line bundle over the massless
orbit.  In four dimensions this bundle has a concrete spinorial
trivialization.  The statements in this and the next two sections concern
tree-level perturbative massless amplitudes after the scattering operator and
LSZ extraction have already been defined; they are not a definition of the
\(S\)-operator.

For a real four-vector \(p^\mu\), define the Hermitian \(2\times2\) matrix
\begin{equation}
  p_{a\dot a}
  =
  p^0\delta_{a\dot a}+p^i(\sigma_i)_{a\dot a},
  \qquad
  a,\dot a=1,2.
  \label{eq:spinor-helicity-momentum-matrix}
\end{equation}
With the mostly-plus metric used in this monograph,
\[
  \det(p_{a\dot a})=(p^0)^2-\vec p^{\,2}=-p^2 .
\]
Thus a complexified null momentum is equivalently a rank-one matrix, and may
be written
\begin{equation}
  p_{a\dot a}
  =
  \lambda_a\widetilde\lambda_{\dot a}.
  \label{eq:null-momentum-spinor-factorization}
\end{equation}
For a real future-directed null momentum one may choose
\(\widetilde\lambda_{\dot a}=\overline{\lambda_a}\), up to a phase.  In
analytic amplitude formulae \(\lambda\) and \(\widetilde\lambda\) are treated
as independent complex spinors; the real physical region is a boundary value.
The factorization \eqref{eq:null-momentum-spinor-factorization} is unchanged
by the little-group scaling
\begin{equation}
  \lambda_a\longmapsto t\,\lambda_a,
  \qquad
  \widetilde\lambda_{\dot a}\longmapsto t^{-1}
  \widetilde\lambda_{\dot a},
  \qquad t\in\mathbb C^\times .
  \label{eq:spinor-helicity-little-group-scaling}
\end{equation}
This is the complexified version of the helicity-frame freedom.

\begin{definition}[Spinor-helicity frame]
\label{def:spinor-helicity-frame}
  A spinor-helicity frame for a complex null momentum \(p\) is a factorization
  \(p_{a\dot a}=\lambda_a\widetilde\lambda_{\dot a}\), with the equivalence
  relation
  \[
    (\lambda,\widetilde\lambda)
    \sim
    (t\lambda,t^{-1}\widetilde\lambda),\qquad t\in\mathbb C^\times .
  \]
  A helicity-\(h\) amplitude component is a section of the corresponding
  helicity line and therefore has weight \(t^{-2h}\) under this scaling.
\end{definition}

Spinor contractions are
\begin{equation}
  \langle ij\rangle
  =
  \epsilon^{ab}\lambda_{i,a}\lambda_{j,b},
  \qquad
  [ij]
  =
  \epsilon^{\dot a\dot b}
  \widetilde\lambda_{i,\dot a}
  \widetilde\lambda_{j,\dot b},
  \label{eq:spinor-helicity-brackets}
\end{equation}
with the same \(\epsilon^{12}=1\) convention as
Section~\ref{sec:spinor-conventions}.  They are antisymmetric:
\(\langle ij\rangle=-\langle ji\rangle\) and \([ij]=-[ji]\).  A direct
determinant calculation gives the mostly-plus invariant
\begin{equation}
  \langle ij\rangle[ji]
  =
  2p_i\cdot p_j.
  \label{eq:spinor-helicity-dot-product}
\end{equation}
For a sum \(P^\mu=\sum_{r\in I}p_r^\mu\), use the abbreviation
\[
  \langle i|P|j]
  :=
  \lambda_i^a P_{a\dot a}\widetilde\lambda_j^{\dot a}.
\]
Momentum conservation for an all-outgoing amplitude is
\begin{equation}
  \sum_{i=1}^n
  \lambda_{i,a}\widetilde\lambda_{i,\dot a}=0 .
  \label{eq:spinor-helicity-momentum-conservation}
\end{equation}

An \(n\)-point helicity amplitude with external helicities \(h_i\) is a
section of the tensor product of the helicity lines.  In a spinor-helicity
frame this means
\begin{equation}
  A_n(\ldots,t_i\lambda_i,t_i^{-1}\widetilde\lambda_i,\ldots)
  =
  t_i^{-2h_i}
  A_n(\ldots,\lambda_i,\widetilde\lambda_i,\ldots).
  \label{eq:spinor-helicity-amplitude-weight}
\end{equation}
This equation is not a mnemonic: it is the helicity character
\(\chi_h\) of the little group, written in the coordinates
\eqref{eq:null-momentum-spinor-factorization}.

\begin{proposition}[Null factorization and mostly-plus spinor invariant]
\label{prop:spinor-helicity-null-factorization-invariant}
  The matrix \(p_{a\dot a}\) has determinant \(-p^2\).  Hence
  \(p^2=0\) is equivalent, over the complex numbers, to the existence of a
  rank-one factorization
  \(p_{a\dot a}=\lambda_a\widetilde\lambda_{\dot a}\).  For two null momenta
  \(p_i=\lambda_i\widetilde\lambda_i\) and
  \(p_j=\lambda_j\widetilde\lambda_j\),
  \[
    \langle ij\rangle[ji]=2p_i\cdot p_j
  \]
  in the mostly-plus convention of this monograph.
\end{proposition}

\begin{proof}
The determinant of \(p^0\mathbf 1+p^i\sigma_i\) is
\((p^0)^2-\vec p^{\,2}=-p^2\).  A \(2\times2\) complex matrix has determinant
zero iff it has rank at most one; the zero matrix corresponds to \(p=0\), and
a nonzero rank-one matrix factors as a column spinor times a row spinor.  For
two rank-one matrices, the antisymmetric contraction identity for \(2\times2\)
matrices gives
\[
  \det(p_i+p_j)
  =
  \epsilon^{ab}\lambda_{i,a}\lambda_{j,b}\,
  \epsilon^{\dot a\dot b}
  \widetilde\lambda_{j,\dot a}\widetilde\lambda_{i,\dot b}
  =
  -\langle ij\rangle[ji].
\]
Since \(p_i^2=p_j^2=0\), the determinant side is
\(-(p_i+p_j)^2=-2p_i\cdot p_j\), yielding
\(\langle ij\rangle[ji]=2p_i\cdot p_j\).
\end{proof}

The vector polarization representatives can also be written spinorially.  Let
\(r_a,\widetilde r_{\dot a}\) be reference spinors with
\(\langle r\,\lambda\rangle\ne0\) and
\([\widetilde\lambda\,\widetilde r]\ne0\).  Define
\begin{equation}
  \varepsilon^+_{a\dot a}(k;r)
  =
  \sqrt2\,\frac{r_a\widetilde\lambda_{\dot a}}
  {\langle r\,\lambda\rangle},
  \qquad
  \varepsilon^-_{a\dot a}(k;r)
  =
  \sqrt2\,\frac{\lambda_a\widetilde r_{\dot a}}
  {[\widetilde\lambda\,\widetilde r]} .
  \label{eq:spinor-helicity-polarization-vectors}
\end{equation}
They obey \(k^{a\dot a}\varepsilon^\pm_{a\dot a}=0\).  Changing the reference
spinors changes \(\varepsilon^\pm\) by a term proportional to
\(k_{a\dot a}=\lambda_a\widetilde\lambda_{\dot a}\), hence by exactly the
representative shift discussed above.  Gauge-invariant amplitudes are
independent of this reference choice.

\section{Color-Ordered Yang--Mills Tree Amplitudes}
\label{sec:color-ordered-tree-amplitudes}

Now specialize to the perturbative tree amplitudes of four-dimensional
Yang--Mills theory, with all external momenta outgoing and the overall
momentum-conservation delta function removed.  The complete color-dressed
amplitude is a sum over trace structures and permutations.  The partial
amplitude
\[
  A_n(1^{h_1},2^{h_2},\ldots,n^{h_n})
\]
denotes the coefficient of a fixed cyclic ordering after stripping the common
coupling and color factor.  The cyclic ordering is part of the datum; poles of
\(A_n\) occur only in sums of adjacent momenta in that cyclic order.

At three points, real massless momenta with momentum conservation are
collinear, but complexified on-shell kinematics has two branches.  On the
holomorphic branch all square brackets vanish and angle brackets need not; on
the antiholomorphic branch all angle brackets vanish and square brackets need
not.  Locality, mass dimension, and the little-group weights
\eqref{eq:spinor-helicity-amplitude-weight} fix the nonzero color-ordered
three-gluon amplitudes, in the normalization used here, to be
\begin{equation}
  A_3(1^-,2^-,3^+)
  =
  \frac{\langle12\rangle^3}
       {\langle23\rangle\langle31\rangle},
  \qquad
  A_3(1^+,2^+,3^-)
  =
  \frac{[12]^3}{[23][31]}.
  \label{eq:three-gluon-spinor-helicity-amplitudes}
\end{equation}
For example, the first formula transforms as
\[
  A_3(1^-,2^-,3^+)
  \longmapsto
  t_1^2t_2^2t_3^{-2}\,
  A_3(1^-,2^-,3^+),
\]
which is exactly the helicity weight for \((-1,-1,+1)\).  Its mass dimension is
that of a tree-level three-gluon partial amplitude after the coupling has been
stripped.

The maximally helicity-violating, or MHV, tree amplitudes are the amplitudes
with exactly two negative-helicity gluons.  With negative-helicity labels
\(r\) and \(s\), the Parke--Taylor formula is
\begin{equation}
  A_n(1^+,\ldots,r^-,\ldots,s^-,\ldots,n^+)
  =
  \frac{\langle rs\rangle^4}
  {\langle12\rangle\langle23\rangle\cdots
   \langle n1\rangle}.
  \label{eq:parke-taylor-mhv}
\end{equation}
The expression is cyclic, has the correct little-group weight on every leg,
and has only adjacent collinear and multi-particle singularities.  The
recursion theorem below turns these consistency checks into a construction
from factorization and the three-point amplitudes.

\begin{theorem}[Adjacent-negative Parke--Taylor tree amplitude]
\label{thm:parke-taylor-mhv-tree}
  In color-ordered tree-level Yang--Mills theory, assuming the three-point
  amplitudes \eqref{eq:three-gluon-spinor-helicity-amplitudes}, tree
  factorization with the all-outgoing internal-line convention displayed
  below, and a BCFW shift for which the boundary term at infinity vanishes, the
  \(n\)-gluon amplitude with negative-helicity gluons at adjacent labels
  \(1\) and \(2\) is
  \[
    A_n(1^-,2^-,3^+,\ldots,n^+)
    =
    \frac{\langle12\rangle^4}
    {\langle12\rangle\langle23\rangle\cdots\langle n1\rangle}.
  \]
\end{theorem}

\begin{proof}
The formula has the correct helicity weight because legs \(1\) and \(2\)
receive \(t_1^2\) and \(t_2^2\), while every positive-helicity leg receives
\(t_i^{-2}\), exactly as required by
\eqref{eq:spinor-helicity-amplitude-weight}.  The base case is the
three-point amplitude displayed in
\eqref{eq:three-gluon-spinor-helicity-amplitudes}, interpreted on the
holomorphic complex branch.

For the induction step use the \((-,+)\) BCFW shift involving one
negative-helicity leg and one positive-helicity leg:
\[
  \widehat{\widetilde\lambda}_1
  =
  \widetilde\lambda_1+z\widetilde\lambda_n,\qquad
  \widehat\lambda_n=\lambda_n-z\lambda_1 .
\]
The boundary term in Theorem~\ref{thm:bcfw-recursion-with-boundary} vanishes
by hypothesis.  The helicity selection rules of the three-point amplitudes
imply that all recursive channels vanish except the channel
\[
  A_{n-1}(\widehat1^-,2^-,3^+,\ldots,(n-2)^+,-\widehat P^+)
  \frac1{P_{1\cdots n-2}^2}
  A_3(\widehat P^-,(n-1)^+,\widehat n^+).
\]
At the pole,
\[
  0=(p_{n-1}+\widehat p_n)^2
   =
   \langle n-1,\widehat n\rangle[\widehat n,n-1],
\]
and the anti-MHV three-point factor is supported on the branch
\(\langle n-1,\widehat n\rangle=0\).  Hence
\[
  z_*=\frac{\langle n-1,n\rangle}{\langle n-1,1\rangle},
  \qquad
  \widehat\lambda_n(z_*)
  =
  \frac{\langle n1\rangle}{\langle n-1,1\rangle}\lambda_{n-1},
\]
where the second identity is Schouten's identity applied to
\((\lambda_{n-1},\lambda_n,\lambda_1)\).  Use the internal spinor frame for
the left momentum
\[
  -\widehat P
  =
  \lambda_{n-1}
  \left(
    \widetilde\lambda_{n-1}
    +\frac{\langle n1\rangle}{\langle n-1,1\rangle}
     \widetilde\lambda_n
  \right).
\]
The induction hypothesis gives
\[
  A_{n-1}
  =
  \frac{\langle12\rangle^4}
  {\langle12\rangle\langle23\rangle\cdots
   \langle n-2,n-1\rangle\langle n-1,1\rangle}.
\]
With the all-outgoing internal-line convention in the factorization residue,
the three-point amplitude and the propagator combine to
\[
  \frac1{P_{1\cdots n-2}^2}
  A_3(\widehat P^-,(n-1)^+,\widehat n^+)
  =
  \frac{\langle n-1,1\rangle}
       {\langle n-1,n\rangle\langle n1\rangle}.
\]
Indeed \(P_{1\cdots n-2}^2=\langle n-1,n\rangle[n,n-1]\), while the
three-point formula evaluates to the displayed numerator factor times
\([n,n-1]\); the relative sign is fixed by assigning the two internal momenta
\(\widehat P\) and \(-\widehat P\) the opposite all-outgoing orientations in the
factorization formula.  Multiplying the last two displayed equations cancels
\(\langle n-1,1\rangle\) and gives the claimed cyclic denominator with the
same numerator \(\langle12\rangle^4\).
\end{proof}

\section{BCFW Recursion}
\label{sec:bcfw-recursion}

Choose two external labels \(i\) and \(j\).  The BCFW deformation
\begin{equation}
  \widehat{\widetilde\lambda}_i(z)
  =
  \widetilde\lambda_i+z\widetilde\lambda_j,
  \qquad
  \widehat\lambda_j(z)
  =
  \lambda_j-z\lambda_i,
  \label{eq:bcfw-shift}
\end{equation}
leaves all other spinors fixed.  It preserves
\(\widehat p_i(z)^2=\widehat p_j(z)^2=0\) and total momentum:
\[
  \widehat p_i(z)+\widehat p_j(z)=p_i+p_j .
\]
Let \(q_{a\dot a}=\lambda_{i,a}\widetilde\lambda_{j,\dot a}\).  Then
\(\widehat p_i=p_i+zq\), \(\widehat p_j=p_j-zq\), and
\(q^2=q\cdot p_i=q\cdot p_j=0\).

\begin{proposition}[BCFW shift preserves the on-shell kinematic surface]
\label{prop:bcfw-shift-on-shell}
  The deformation \eqref{eq:bcfw-shift} preserves the null conditions of legs
  \(i\) and \(j\), preserves total momentum, and is generated by a null vector
  \(q_{a\dot a}=\lambda_{i,a}\widetilde\lambda_{j,\dot a}\) satisfying
  \(q^2=q\cdot p_i=q\cdot p_j=0\).
\end{proposition}

\begin{proof}
The shifted momenta are
\[
  \widehat p_i(z)=
  \lambda_i(\widetilde\lambda_i+z\widetilde\lambda_j)=p_i+zq,
  \qquad
  \widehat p_j(z)=
  (\lambda_j-z\lambda_i)\widetilde\lambda_j=p_j-zq.
\]
Each remains rank one, hence null.  Their sum is unchanged.  The vector
\(q=\lambda_i\widetilde\lambda_j\) is rank one and therefore null; moreover
\(\langle ii\rangle=0\) and \([jj]=0\), so
\(2q\cdot p_i=2q\cdot p_j=0\), with the mostly-plus sign understood through
Proposition~\ref{prop:spinor-helicity-null-factorization-invariant}.
\end{proof}

At tree level a color-ordered amplitude \(A_n(z)\) is a rational function of
\(z\).  Its finite poles occur when a cyclically adjacent momentum sum
\[
  P_I=\sum_{\ell\in I}p_\ell
\]
contains exactly one of the shifted labels and becomes null:
\begin{equation}
  \widehat P_I(z_I)^2=0.
  \label{eq:bcfw-factorization-pole}
\end{equation}
Near such a pole, tree-level unitarity and locality give the factorization
\begin{equation}
  A_n(z)
  \sim
  \sum_{h}
  A_L(z_I)\,
  \frac{1}{\widehat P_I(z)^2}\,
  A_R(z_I),
  \label{eq:tree-factorization-bcfw}
\end{equation}
where the two lower-point amplitudes have an internal on-shell gluon of
helicity \(h\) and \(-h\), respectively.  Applying Cauchy's theorem to
\(A_n(z)/z\) gives
\begin{equation}
  A_n(0)
  =
  \sum_{I,h}
  A_L(z_I)\,
  \frac{1}{P_I^2}\,
  A_R(z_I)
  +B_\infty .
  \label{eq:bcfw-recursion-with-boundary}
\end{equation}
The boundary term \(B_\infty\) is the residue at \(z=\infty\).  For
Yang--Mills tree amplitudes it vanishes for the standard good shifts used in
the MHV construction, including the \((-,+)\) shift below.  When a shift has
nonzero boundary term, equation \eqref{eq:bcfw-recursion-with-boundary} remains
true but is not a closed recursion until \(B_\infty\) is determined by other
input.

\begin{theorem}[BCFW recursion with boundary term]
\label{thm:bcfw-recursion-with-boundary}
  Let \(A_n(z)\) be a tree-level color-ordered Yang--Mills partial amplitude
  under a BCFW shift.  Suppose its only finite singularities are simple
  factorization poles where an adjacent shifted momentum sum becomes null, and
  suppose the residues are the tree factorization products
  \[
    \operatorname*{Res}_{z=z_I} A_n(z)
    =
    \sum_h
    A_L(z_I)\,
    \operatorname*{Res}_{z=z_I}\!\left(\frac1{\widehat P_I(z)^2}\right)
    A_R(z_I).
  \]
  Then
  \[
    A_n(0)
    =
    \sum_{I,h}A_L(z_I)\frac1{P_I^2}A_R(z_I)+B_\infty,
  \]
  where \(B_\infty\) is minus the residue of \(A_n(z)/z\) at infinity.  If
  \(A_n(z)\to0\) as \(z\to\infty\), then \(B_\infty=0\).
\end{theorem}

\begin{proof}
The meromorphic one-form \(A_n(z)\,dz/z\) on the Riemann sphere has total
residue zero.  Its residue at \(z=0\) is \(A_n(0)\).  A finite pole occurs
when \(\widehat P_I(z)^2=P_I^2+z\,2P_I\cdot q\) vanishes, so
\[
  z_I=-\frac{P_I^2}{2P_I\cdot q}.
\]
Near this point,
\[
  \frac1{\widehat P_I(z)^2}
  =
  \frac1{(z-z_I)\,2P_I\cdot q}
  =
  \frac{z_I}{(z-z_I)(-P_I^2)} .
\]
Therefore the contribution of this finite pole to the residue of \(A_n(z)/z\)
is
\(-A_L(z_I)A_R(z_I)/P_I^2\), summed over internal helicities.  Moving all
finite-pole residues and the residue at infinity to the other side gives the
displayed recursion formula.
\end{proof}

Figure~\ref{fig:bcfw-factorization} records the finite-pole factorization
geometry used in the recursion formula: the shifted amplitude splits into left
and right on-shell subamplitudes joined by the channel \(\widehat P_I^2=0\).

\begin{figure}[H]
  \centering
  \begin{tikzpicture}[>=Stealth,scale=0.94,
    blob/.style={ellipse,draw=black,thick,minimum width=1.65cm,
      minimum height=1.05cm,fill=blue!5}]
    \node[blob] (L) at (-2.2,0) {\(A_L(z_I)\)};
    \node[blob] (R) at (2.2,0) {\(A_R(z_I)\)};
    \draw[thick] (L.east)--node[above] {\(\widehat P_I^2=0\)} (R.west);
    \draw[thick,->] (-4.0,0.85)--(L.west);
    \draw[thick,->] (-4.0,0.15)--(L.west);
    \draw[thick,->] (-4.0,-0.65)--(L.west);
    \draw[thick,->] (R.east)--(4.0,0.65);
    \draw[thick,->] (R.east)--(4.0,-0.15);
    \draw[thick,->] (R.east)--(4.0,-0.85);
    \node[left] at (-4.0,0.85) {\(\hat i\)};
    \node[right] at (4.0,0.65) {\(\hat j\)};
    \node[align=center] at (0,-1.25)
      {BCFW pole = ordinary tree factorization\\
       with complex on-shell shifted momenta};
  \end{tikzpicture}
  \caption{A pole in the complex BCFW parameter occurs when an adjacent
  momentum sum becomes null.  The residue is the product of two lower-point
  on-shell amplitudes and the intermediate propagator.}
  \label{fig:bcfw-factorization}
\end{figure}

\section{Parke--Taylor from Recursion}
\label{sec:parke-taylor-from-bcfw}

Consider
\[
  A_n(1^-,2^-,3^+,\ldots,n^+).
\]
Use the shift
\[
  \widehat{\widetilde\lambda}_1
  =
  \widetilde\lambda_1+z\widetilde\lambda_n,
  \qquad
  \widehat\lambda_n
  =
  \lambda_n-z\lambda_1 .
  \label{eq:mhv-bcfw-shift}
\]
This is a good \((-,+)\) Yang--Mills shift, so \(B_\infty=0\).  The BCFW
terms in which one side has fewer than two negative helicities vanish, except
for the three-point anti-MHV factor allowed by complex kinematics.  Inductively
one obtains a single nonzero channel, consisting of an \((n-1)\)-point MHV
amplitude and a three-point amplitude.  Substitution of
\eqref{eq:three-gluon-spinor-helicity-amplitudes} and the induction hypothesis
gives
\[
  A_n(1^-,2^-,3^+,\ldots,n^+)
  =
  \frac{\langle12\rangle^4}
  {\langle12\rangle\langle23\rangle\cdots\langle n1\rangle}.
\]
This proves \eqref{eq:parke-taylor-mhv} for this placement of the two negative
helicities.  The nonadjacent \((r,s)\) formula follows from the same BCFW
recursion together with Schouten identities, but more than one channel can
contribute; the adjacent proof above is the single-channel case needed for the
worked example below.

For the first genuinely five-leg case,
\[
  A_5(1^-,2^-,3^+,4^+,5^+),
\]
the shift \eqref{eq:mhv-bcfw-shift} gives one nonzero recursive contribution:
\begin{equation}
  A_5
  =
  A_4(\widehat{1}^-,2^-,3^+,-\widehat P^+)\,
  \frac{1}{P_{123}^2}\,
  A_3(\widehat P^-,4^+,\widehat{5}^+),
  \label{eq:five-point-mhv-bcfw-channel}
\end{equation}
where \(P_{123}=p_1+p_2+p_3\) and
\(\widehat P=\widehat p_1(z_*)+p_2+p_3\) is null.  The pole is determined by
\[
  \widehat P(z_*)^2=0.
\]
Using the four-point Parke--Taylor amplitude on the left and the
three-point amplitude \eqref{eq:three-gluon-spinor-helicity-amplitudes} on the
right, all dependence on the internal spinor representation of
\(\widehat P\) cancels.  The result is
\begin{equation}
  A_5(1^-,2^-,3^+,4^+,5^+)
  =
  \frac{\langle12\rangle^4}
  {\langle12\rangle\langle23\rangle\langle34\rangle
   \langle45\rangle\langle51\rangle}.
  \label{eq:five-point-parke-taylor}
\end{equation}
The calculation illustrates the conceptual point of BCFW recursion: tree
amplitudes are determined by their on-shell factorization residues, helicity
weights, and large-complex-momentum behavior, without introducing off-shell
gauge-dependent subamplitudes.

The next chapter supplies the dynamics whose one-particle sector realizes
these helicity states: Maxwell theory, its constraints, and the controlled
introduction of gauge fixing.
