\chapter{Massive Particles and Spin}
\label{ch:massive-particles-spin}
\ConventionsLorentzian


\section{Massive One-Particle Representations}

\begin{definition}[Massive one-particle spin datum]
\label{def:massive-one-particle-spin-datum}
Work in \(D=4\). A massive one-particle datum consists of a Hilbert space
\(\Hilb_1\) carrying an irreducible positive-energy unitary representation of
the Poincare group whose translation generators \(P^\mu\) have joint spectrum
supported on a single positive-energy mass shell of mass \(m>0\).
Equivalently, it is the massive row of
Table~\ref{tab:wigner-classification-four-dimensional}. The mass shell is
\[
  \Sigma_m^+
  =
  \{p=(p^0,\vec p):p^0=\omega_{\vec p},\ p^2=-m^2\},
  \qquad
  \omega_{\vec p}=\sqrt{\vec p^{\,2}+m^2}.
\]
Generalized momentum states, understood as distributional vectors in the
rigged-Hilbert-space sense of
Section~\ref{sec:rigged-hilbert-continuous-spectrum}, are normalized by
\[
  \langle \vec q,\tau|\vec p,\sigma\rangle
  =
  (2\pi)^3\,2\omega_{\vec p}\,
  \delta^{(3)}(\vec q-\vec p)\,\delta_{\tau\sigma}.
\]
The labels \(\sigma,\tau\) span a finite-dimensional internal vector space
which is fixed by Lorentz symmetry.
\end{definition}

\begin{remark}[Spacelike Wigner orbits]
The full representation theory of the Poincare group also contains spacelike
orbits \(p^2=\mu^2>0\), sometimes called tachyonic orbits.  They are absent
from this chapter and the following massless chapter because the spectrum
condition assumed in Chapter~\ref{ch:relativistic-qm-local-operator-structure}
requires \(\operatorname{Spec}(P)\subset\overline V_+\), while a spacelike orbit
lies outside the closed forward lightcone.
\end{remark}

\begin{definition}[Standard rest momentum and normalization frames]
\label{def:massive-standard-rest-momentum-normalization}
Choose the reference momentum
\[
  p_R=(m,\vec0).
\]
For every \(p\in\Sigma_m^+\), choose a standard Lorentz transformation
\[
  L(p)\in SO^+(1,3),
  \qquad
  L(p)p_R=p,
  \qquad
  L(p_R)=1 .
\]
It is sometimes useful to compare this covariant normalization with the
delta-function normalization
\[
  {}_\delta\langle \vec q,\tau|\vec p,\sigma\rangle_\delta
  =
  \delta^{(3)}(\vec q-\vec p)\delta_{\tau\sigma}.
\]
In that convention one may define
\[
  |\vec p,\sigma\rangle_\delta
  =
  N(\vec p)\,U(L(p))|\vec0,\sigma\rangle_\delta ,
  \qquad
  N(\vec p)=\left(\frac{m}{\omega_{\vec p}}\right)^{1/2}.
\]
The factor \(N(\vec p)\) is the Jacobian needed to keep the measure
\(d^3\vec p\) compatible with Lorentz transformations in a
delta-normalized basis. The covariantly normalized states used below absorb
this factor into the invariant mass-shell measure.
\end{definition}

\begin{definition}[Massive Wigner rotation]
\label{def:massive-wigner-rotation}
The stabilizer of \(p_R\) in \(SO^+(1,3)\) is the rotation group \(SO(3)\).
Whenever a Wigner rotation is inserted into a spin representation below, it is
understood through its lift to the universal cover of \(SO(3)\); the reason for
this lift is explained in
Section~\ref{sec:massive-spin-projective-cover}.  The same letter denotes the
projected Lorentz transformation when it acts on four-vectors.

For a Lorentz transformation \(\Lambda\), define the Wigner rotation
\[
  W(\Lambda,p)
  =
  L(\Lambda p)^{-1}\Lambda L(p).
\]
It obeys
\[
  W(\Lambda,p)p_R=p_R,
\]
and is therefore an element of the massive little group. If the particle has
spin \(j\), with
\[
  j=0,\frac12,1,\frac32,\ldots,
\]
let \(D^{(j)}\) be the corresponding irreducible unitary representation of
\(SU(2)\) on a vector space \(V_j\) of dimension \(2j+1\).  For a four-vector
\(r\), write \(\operatorname{sp}(r)\) for its spatial part, so
\[
  p=(\omega_{\vec p},\vec p),
  \qquad
  \operatorname{sp}(p)=\vec p .
\]
\end{definition}

\paragraph{Massive Wigner transformation law.}
The covariantly normalized one-particle states transform as
\[
  U(\Lambda)|\vec p,\sigma\rangle
  =
  \sum_{\tau=-j}^{j}
  D^{(j)}_{\tau\sigma}\!\left(W(\Lambda,p)\right)
  |\operatorname{sp}(\Lambda p),\tau\rangle ,
\]
where \(\sigma,\tau=-j,-j+1,\ldots,j\). The covariant normalization above
absorbs the Jacobian of the mass-shell measure, so no additional
\(\sqrt{(\Lambda p)^0/p^0}\) factor appears.
In the delta-normalized convention the same computation gives
\[
  U(\Lambda)|\vec p,\sigma\rangle_\delta
  =
  \left(\frac{(\Lambda p)^0}{p^0}\right)^{1/2}
  \sum_{\tau=-j}^{j}
  D^{(j)}_{\tau\sigma}\!\left(W(\Lambda,p)\right)
  |\operatorname{sp}(\Lambda p),\tau\rangle_\delta .
\]
This is the same representation written in a basis adapted to
\(d^3\vec p\) rather than to \(d^3\vec p/(2p^0)\). Keeping both formulas
visible prevents the normalization factor from being mistaken for a property
of spin.
By definition \(L(p)p_R=p\) and \(L(\Lambda p)p_R=\Lambda p\). Therefore
\[
  W(\Lambda,p)p_R
  =
  L(\Lambda p)^{-1}\Lambda L(p)p_R
  =
  p_R ,
\]
so \(W(\Lambda,p)\) belongs to the little group of the rest momentum.  The
restriction of the Poincare representation to the little group at \(p_R\) is an
irreducible \(SU(2)\) representation \(D^{(j)}\).  Acting first by \(L(p)\),
then by \(\Lambda\), and finally comparing with the chosen standard
transformation \(L(\Lambda p)\) gives exactly the displayed spin matrix
\(D^{(j)}(W(\Lambda,p))\).

The absence or presence of the square-root factor is a statement about the
basis measure.  The invariant mass-shell measure satisfies
\[
  \frac{d^3\vec p}{2p^0}
  =
  \frac{d^3\operatorname{sp}(\Lambda p)}{2(\Lambda p)^0}.
\]
Equivalently,
\[
  d^3\operatorname{sp}(\Lambda p)
  =
  \frac{(\Lambda p)^0}{p^0}\,d^3\vec p .
\]
Thus the covariantly normalized basis, whose resolution of the identity uses
\(d^3\vec p/(2p^0)\), transforms without a scalar Jacobian.  A basis
normalized against \(d^3\vec p\) must instead carry the unitary half-density
factor \(((\Lambda p)^0/p^0)^{1/2}\).  This proves both formulas.

\section{Projective Rotations and the Covering Group}
\label{sec:massive-spin-projective-cover}

The appearance of \(SU(2)\) rather than only \(SO(3)\) is forced by the
quantum implementation of rotations. A physical pure state is a ray, so the
operator associated with a rotation \(R\in SO(3)\) is defined on Hilbert
space only up to a phase if no continuity convention has been chosen. Thus a
projective representation may obey
\[
  U(R_2)U(R_1)
  =
  \ee^{\ii\alpha(R_2,R_1)}U(R_2R_1)
\]
for a real phase function \(\alpha\). The phase ambiguity is removed by
lifting rotations to paths. If \(\gamma:[0,1]\to SO(3)\) is a continuous path
with \(\gamma(0)=1\) and \(\gamma(1)=R\), continuity of the infinitesimal
generators determines an operator \(U_\gamma\) starting from \(U_{\gamma(0)}=1\).
If \(\gamma\) and \(\gamma'\) are homotopic with endpoints fixed, then
\[
  U_\gamma=U_{\gamma'} .
\]
Consequently the correct group acting linearly is the universal cover of
\(SO(3)\), whose points are endpoint-fixed homotopy classes of such paths.

\begin{center}
\begin{tikzpicture}[scale=0.95,>=Stealth]
  \draw[gray!50,thick] (-3,0) circle (1.15);
  \fill (-4.15,0) circle (1.8pt) node[left] {\(1\)};
  \fill (-1.85,0) circle (1.8pt) node[right] {\(R\)};
  \draw[blue!75!black,very thick,->]
    (-4.05,0.12).. controls (-3.7,0.95) and (-2.3,0.95) .. (-1.95,0.12);
  \draw[blue!75!black,very thick,->]
    (-4.05,-0.12).. controls (-3.7,-0.95) and (-2.3,-0.95) .. (-1.95,-0.12);
  \node[blue!75!black] at (-3,1.55) {homotopic paths};

  \draw[gray!50,thick] (2.3,0) circle (1.15);
  \fill (1.15,0) circle (1.8pt) node[left] {\(1\)};
  \fill (3.45,0) circle (1.8pt) node[right] {\(1\)};
  \draw[red!75!black,very thick,->]
    (1.23,0.02).. controls (1.55,1.08) and (3.05,1.08) .. (3.37,0.02);
  \draw[red!75!black,very thick,->]
    (3.37,-0.02).. controls (3.05,-1.08) and (1.55,-1.08) .. (1.23,-0.02);
  \node[red!75!black] at (2.3,1.55) {\(2\pi\) loop};
  \node[align=center] at (2.3,-1.62)
    {not contractible in \(SO(3)\)\\but its square is};
\end{tikzpicture}
\end{center}

\paragraph{Massive spin labels from the covering group.}
The universal cover of the massive rotation little group is
\[
  \widetilde{SO(3)}\simeq SU(2).
\]
Its finite-dimensional irreducible unitary representations are indexed by
\[
  j=0,\frac12,1,\frac32,\ldots ,
\]
with representation space \(V_j\) of dimension \(2j+1\). The nontrivial
central element of \(SU(2)\), namely the lift of a \(2\pi\) rotation, acts by
\[
  D^{(j)}(-1)=(-1)^{2j}\mathbf 1_{V_j}.
\]
The lift of a \(4\pi\) rotation is the identity.  Hence an irreducible massive
one-particle representation carries exactly one such spin label \(j\).
The group \(SO(3)\) is \(\mathbb{RP}^3\). Its fundamental group is therefore
\[
  \pi_1(SO(3))\simeq \mathbb Z_2 .
\]
The nontrivial class is represented by a \(2\pi\) rotation loop, while the
double loop representing a \(4\pi\) rotation is contractible. The simply
connected compact Lie group with Lie algebra \(\mathfrak{so}(3)\) is
\(SU(2)\), and the covering map \(SU(2)\to SO(3)\) has kernel
\(\{\pm 1\}\).  The highest-weight classification of irreducible unitary
\(SU(2)\) representations gives one representation \(D^{(j)}\) for
each nonnegative half-integer \(j\), of dimension \(2j+1\).  In the
spin-\(j\) representation the central element \(-1\) is the exponential of a
\(2\pi\) rotation and acts by \((-1)^{2j}\).  Irreducibility of the
one-particle Poincare representation means that its little-group
representation is one irreducible summand, not a direct sum.  With column
vectors the matrices obey \(D(W_2W_1)=D(W_2)D(W_1)\); row-index conventions on
bras or matrix elements transpose this order and do not alter the
representation.

\section{Spin Frames on the Mass Shell}

\begin{definition}[Massive spin frame]
\label{def:massive-spin-frame}
The choice of standard transformations \(L(p)\) is a choice of spin frame over
the mass shell. To make this statement precise, let
\[
  d\mu_m(p)=\frac{d^3p}{(2\pi)^3\,2\omega_{\vec p}}
\]
be the invariant measure on \(\Sigma_m^+\). After choosing \(L(p)\), the
one-particle Hilbert space is represented as
\[
  \Hilb_1\simeq L^2(\Sigma_m^+,d\mu_m;V_j).
\]
The generalized basis associated to this frame is the distributional map
\[
  F_L(p):V_j\longrightarrow \Hilb_1,
  \qquad
  F_L(p)e_\sigma=|\vec p,\sigma\rangle ,
\]
where \(\{e_\sigma\}_{\sigma=-j}^j\) is an orthonormal basis of \(V_j\).
\end{definition}

\begin{lemma}[Wigner cocycle and spin-frame change]
\label{lem:massive-wigner-cocycle-spin-frame-change}
The Lorentz action is
\[
  U(\Lambda)F_L(p)
  =
  F_L(\Lambda p)D^{(j)}(W(\Lambda,p)).
\]
This formula is a representation of the Lorentz group because the Wigner
rotation obeys the cocycle identity
\[
  W(\Lambda_2\Lambda_1,p)
  =
  W(\Lambda_2,\Lambda_1 p)W(\Lambda_1,p).
\]
If \(Q:\Sigma_m^+\to SU(2)\) is a smooth little-group-valued function, then
\[
  L'(p)=L(p)Q(p)
\]
is another choice of standard transformations. The corresponding Wigner
rotation is
\[
  W'(\Lambda,p)
  =
  Q(\Lambda p)^{-1}W(\Lambda,p)Q(p).
\]
The associated generalized basis is
\[
  F_{L'}(p)=F_L(p)D^{(j)}(Q(p)).
\]
Thus a wavefunction written as a column \(\psi_L(p)\in V_j\) in the \(L\)
frame is written in the \(L'\) frame as
\[
  \psi_{L'}(p)=D^{(j)}(Q(p))^{-1}\psi_L(p).
\]
The inner product
\[
  \langle \psi,\chi\rangle
  =
  \int_{\Sigma_m^+} d\mu_m(p)\,
  \psi_L(p)^\dagger\chi_L(p)
\]
is independent of the spin frame because \(D^{(j)}(Q(p))\) is unitary. In this
language, a scattering amplitude with external massive spin states is a
multilinear pairing of spin-frame components and dual spin-frame components
whose value is invariant under the simultaneous transformations above.
\end{lemma}

\begin{proof}
The action formula is the massive Wigner transformation law above written as a
map \(V_j\to\Hilb_1\). The cocycle identity follows from a direct
factorization:
\[
\begin{aligned}
  W(\Lambda_2\Lambda_1,p)
  &=
  L(\Lambda_2\Lambda_1p)^{-1}\Lambda_2\Lambda_1L(p)\\
  &=
  L(\Lambda_2\Lambda_1p)^{-1}\Lambda_2L(\Lambda_1p)\,
  L(\Lambda_1p)^{-1}\Lambda_1L(p)\\
  &=
  W(\Lambda_2,\Lambda_1p)W(\Lambda_1,p).
\end{aligned}
\]
If \(L'(p)=L(p)Q(p)\), then
\[
\begin{aligned}
  W'(\Lambda,p)
  &=
  L'(\Lambda p)^{-1}\Lambda L'(p)\\
  &=
  Q(\Lambda p)^{-1}L(\Lambda p)^{-1}\Lambda L(p)Q(p)\\
  &=
  Q(\Lambda p)^{-1}W(\Lambda,p)Q(p).
\end{aligned}
\]
The basis relation
\[
  F_{L'}(p)=F_L(p)D^{(j)}(Q(p))
\]
therefore forces the component relation
\(\psi_{L'}(p)=D^{(j)}(Q(p))^{-1}\psi_L(p)\).  Since \(D^{(j)}\) is unitary,
\[
  \psi_{L'}(p)^\dagger\chi_{L'}(p)
  =
  \psi_L(p)^\dagger\chi_L(p)
\]
pointwise on the mass shell, and the displayed inner product is
spin-frame independent.  Dual spin-frame components transform by the inverse
dual matrices, so any contracted scattering amplitude is independent of the
choice of \(L(p)\).
\end{proof}

\section{Covariant Fields as Intertwiners}

\begin{proposition}[Vacuum-to-particle intertwiners]
\label{prop:covariant-field-intertwiner-equivariance}
Let \(\widehat\Phi_A(x)\) be a local field whose index \(A\) transforms in a
finite-dimensional representation \(R\) of the double cover
\(\operatorname{Spin}^+(1,3)\simeq SL(2,\mathbb C)\):
\[
  U(\Lambda)\widehat\Phi_A(x)U(\Lambda)^{-1}
  =
  R_A{}^B(\Lambda)\widehat\Phi_B(\Lambda x).
\]
The vacuum-to-particle matrix element is
\[
  \bra{\vac}\widehat\Phi_A(0)|\vec p,\sigma\rangle
  =
  u_A{}^\sigma(p).
\]
Translation covariance gives the spacetime dependence
\[
  \bra{\vac}\widehat\Phi_A(x)|\vec p,\sigma\rangle
  =
  \ee^{\ii p\cdot x}u_A{}^\sigma(p).
\]
Lorentz covariance and the Wigner transformation law imply
\[
  \sum_{\tau=-j}^j
  u_A{}^\tau(\Lambda p)
  D^{(j)}_{\tau\sigma}(W(\Lambda,p))
  =
  R_A{}^B(\Lambda^{-1})u_B{}^\sigma(p).
\]
This is the equivariance condition that connects the covariant field index
\(A\) to the spin-frame index \(\sigma\).

At the reference momentum \(p_R\), the matrix
\[
  u_A{}^\sigma(p_R):V_j\longrightarrow R|_{SU(2)}
\]
is an intertwiner of \(SU(2)\) representations:
\[
  R(W)_A{}^B u_B{}^\sigma(p_R)
  =
  \sum_{\tau=-j}^j
  u_A{}^\tau(p_R)D^{(j)}_{\tau\sigma}(W),
  \qquad W\in SU(2).
\]
Thus a covariant field can have nonzero overlap with a spin-\(j\) particle
exactly through the spin-\(j\) part contained in the restriction of \(R\) to
the massive little group. The boosted wavefunctions are
\[
  u_A{}^\sigma(p)
  =
  R_A{}^B(L(p)^{-1})\,u_B{}^\sigma(p_R),
\]
with the convention for \(L(p)\) fixed above. Different choices of standard
boost are spin-frame changes of the type described in the previous section,
and the field intertwiners are used with the corresponding transformed
spin-frame components.
\end{proposition}

\begin{proof}
Translation covariance gives
\[
  \widehat\Phi_A(x)=\ee^{-\ii P\cdot x}\widehat\Phi_A(0)\ee^{\ii P\cdot x}
\]
in the convention used throughout this volume.  Since
\(P^\mu|\vec p,\sigma\rangle=p^\mu|\vec p,\sigma\rangle\) and
\(\bra{\vac}P^\mu=0\), the spacetime dependence is
\(\ee^{\ii p\cdot x}u_A{}^\sigma(p)\).

For Lorentz covariance, compute the same matrix element in two ways.  The
Wigner transformation law gives
\[
  \bra{\vac}\widehat\Phi_A(0)U(\Lambda)|\vec p,\sigma\rangle
  =
  \sum_{\tau=-j}^{j}
  u_A{}^\tau(\Lambda p)D^{(j)}_{\tau\sigma}(W(\Lambda,p)).
\]
Vacuum invariance and field covariance give
\[
\begin{aligned}
  \bra{\vac}\widehat\Phi_A(0)U(\Lambda)|\vec p,\sigma\rangle
  &=
  \bra{\vac}U(\Lambda)
  U(\Lambda)^{-1}\widehat\Phi_A(0)U(\Lambda)|\vec p,\sigma\rangle\\
  &=
  R_A{}^B(\Lambda^{-1})u_B{}^\sigma(p).
\end{aligned}
\]
Equating these expressions proves the equivariance condition.  If
\(\Lambda=W\) fixes \(p_R\), and the standard section has \(L(p_R)=1\), the
equivariance equation becomes precisely the displayed \(SU(2)\)-intertwiner
condition at the rest momentum.  Conversely, a rest intertwiner satisfying this
condition determines the boosted functions by
\[
  u_A{}^\sigma(p)=R_A{}^B(L(p)^{-1})u_B{}^\sigma(p_R),
\]
and substituting the cocycle identity of
Lemma~\ref{lem:massive-wigner-cocycle-spin-frame-change} verifies the
equivariance equation for arbitrary \(\Lambda\).  The possible nonzero
intertwiners are therefore exactly the copies of \(V_j\) appearing in the
restriction of \(R\) to the massive little group.
\end{proof}

\section{Spin One Half}

For spin \(j=\frac12\), the smallest Lorentz-covariant field space is a Dirac
spinor. Let the gamma matrices obey
\[
  \{\gamma^\mu,\gamma^\nu\}=2\eta^{\mu\nu},
  \qquad
  \eta=\operatorname{diag}(-,+,+,+).
\]
The global spinor and gamma-matrix conventions are collected in
Section~\ref{sec:spinor-conventions}; the present section uses the same
symbols and fixes the particle-intertwiner normalizations.
One concrete realization, useful for checking signs, is
\[
  \gamma^0
  =
  -\ii
  \begin{pmatrix}
    0&I_2\\ I_2&0
  \end{pmatrix},
  \qquad
  \gamma^i
  =
  -\ii
  \begin{pmatrix}
    0&\sigma_i\\ -\sigma_i&0
  \end{pmatrix},
  \qquad i=1,2,3,
\]
where
\[
  \sigma_1=\begin{pmatrix}0&1\\1&0\end{pmatrix},
  \quad
  \sigma_2=\begin{pmatrix}0&-\ii\\ \ii&0\end{pmatrix},
  \quad
  \sigma_3=\begin{pmatrix}1&0\\0&-1\end{pmatrix}.
\]
The following formulas are basis-independent statements about the Clifford
module.
Define
\[
  S^{\mu\nu}
  =
  -\frac{\ii}{4}[\gamma^\mu,\gamma^\nu],
  \qquad
  R(\Lambda)
  =
  \exp\!\left(-\frac{\ii}{2}\omega_{\mu\nu}S^{\mu\nu}\right)
\]
for
\[
  \Lambda=\exp(\omega),
  \qquad
  \omega_{\mu\nu}=-\omega_{\nu\mu}.
\]
The generators satisfy
\[
  [S^{\mu\nu},\gamma^\rho]
  =
  -\ii\gamma^\mu\eta^{\nu\rho}
  +\ii\gamma^\nu\eta^{\mu\rho},
\]
and therefore
\[
  R(\Lambda)\gamma^\mu R(\Lambda)^{-1}
  =
  \Lambda^\mu{}_\nu\gamma^\nu .
\]
On the rest subspace, a spatial rotation by angle \(\theta\) about the
\(3\)-axis acts in the spin-\(\frac12\) representation as
\[
  D\!\left(R_{12}(\theta)\right)
  =
  \exp\!\left(-\frac{\ii\theta}{2}\sigma_3\right)
  =
  \begin{pmatrix}
    \ee^{-\ii\theta/2}&0\\
    0&\ee^{\ii\theta/2}
  \end{pmatrix}.
\]
Thus a \(2\pi\) rotation acts by \(-1\) on a spinor, while a \(4\pi\)
rotation acts by \(+1\). This is the infinitesimal form of the covering-group
discussion above.

\begin{proposition}[Massive Dirac eigenspaces, projectors, and spin sums]
\label{prop:massive-dirac-projectors-spin-sums}
Write
\[
  \not p=p_\mu\gamma^\mu .
\]
For \(p_R=(m,\vec0)\),
\[
  \not p_R=-m\gamma^0,
  \qquad
  \not p_R^{\,2}=-m^2 .
\]
Choose two rest spinors \(u_R^\sigma\) and two rest spinors \(v_R^\sigma\),
\(\sigma=1,2\), by
\[
  \not p_R u_R^\sigma=\ii m\,u_R^\sigma,
  \qquad
  \not p_R v_R^\sigma=-\ii m\,v_R^\sigma .
\]
Their boosts are
\[
  u^\sigma(p)=R(L(p)^{-1})u_R^\sigma,
  \qquad
  v^\sigma(p)=R(L(p)^{-1})v_R^\sigma .
\]
Using covariance of \(\gamma^\mu\), these spinors obey
\[
  (\not p-\ii m)u^\sigma(p)=0,
  \qquad
  (\not p+\ii m)v^\sigma(p)=0.
\]
These are the momentum-space Dirac equations associated with the two
eigenspaces of \(\not p\) on the mass shell.

Let
\[
  \beta=\ii\gamma^0,
  \qquad
  \bar u=u^\dagger\beta,
  \qquad
  \bar v=v^\dagger\beta .
\]
The hermiticity identities in the above signature are
\[
  (\gamma^0)^\dagger=-\gamma^0,
  \qquad
  (\gamma^i)^\dagger=\gamma^i,
  \qquad
  R(\Lambda)^\dagger=\beta R(\Lambda^{-1})\beta .
\]
They imply that the \(\beta\)-pairing is Lorentz invariant:
\[
  (R(\Lambda)s)^\dagger\beta R(\Lambda)t=s^\dagger\beta t .
\]
The projectors onto the two eigenspaces of \(\not p\) are
\[
  \Pi_+(p)=\frac{m-\ii\not p}{2m},
  \qquad
  \Pi_-(p)=\frac{m+\ii\not p}{2m}.
\]
For the delta-normalized one-particle states it is useful to include the same
Jacobian factor as in the state normalization and define
\[
  \mathcal U^\sigma(p)
  =
  \left(\frac{m}{p^0}\right)^{1/2}R(L(p)^{-1})\mathcal U_R^\sigma,
  \qquad
  \mathcal V^\sigma(p)
  =
  \left(\frac{m}{p^0}\right)^{1/2}R(L(p)^{-1})\mathcal V_R^\sigma .
\]
Choose rest spinors so that
\[
  \mathcal U_R^{\sigma\dagger}\beta\mathcal U_R^\tau=\delta^{\sigma\tau},
  \qquad
  \mathcal V_R^{\sigma\dagger}\beta\mathcal V_R^\tau=-\delta^{\sigma\tau},
  \qquad
  \mathcal U_R^{\sigma\dagger}\beta\mathcal V_R^\tau=0 .
\]
Then
\[
  \mathcal U^\sigma(p)^\dagger\beta\mathcal U^\tau(p)
  =
  \frac{m}{p^0}\delta^{\sigma\tau},
  \qquad
  \mathcal V^\sigma(p)^\dagger\beta\mathcal V^\tau(p)
  =
  -\frac{m}{p^0}\delta^{\sigma\tau}.
\]
The corresponding spin sums, with ordinary matrix multiplication in spinor
indices, are
\[
  \sum_{\sigma=1}^2
  \mathcal U^\sigma(p)\mathcal U^\sigma(p)^\dagger
  =
  \frac{m-\ii\not p}{2p^0}\,\beta,
  \qquad
  \sum_{\sigma=1}^2
  \mathcal V^\sigma(p)\mathcal V^\sigma(p)^\dagger
  =
  \frac{-m-\ii\not p}{2p^0}\,\beta .
\]
Equivalently,
\[
  \sum_{\sigma=1}^2
  \mathcal U^\sigma(p)\overline{\mathcal U}^{\,\sigma}(p)
  =
  \frac{m-\ii\not p}{2p^0},
  \qquad
  \sum_{\sigma=1}^2
  \mathcal V^\sigma(p)\overline{\mathcal V}^{\,\sigma}(p)
  =
  -\frac{m+\ii\not p}{2p^0}.
\]
The minus sign in the second identity is fixed by the sign of the
\(\beta\)-norm on the negative Dirac eigenspace; the particle Hilbert-space
inner product remains positive.
\end{proposition}

\begin{proof}
The Clifford relation gives
\[
  \not p^{\,2}
  =
  p_\mu p_\nu\gamma^\mu\gamma^\nu
  =
  p^2\mathbf 1
  =
  -m^2\mathbf 1
\]
on the mass shell, because the antisymmetric commutator part drops out after
contraction with \(p_\mu p_\nu\). Hence the minimal polynomial of \(\not p\)
divides \(\lambda^2+m^2\), and the two spectral projectors are
\[
  \Pi_+(p)=\frac{m-\ii\not p}{2m},
  \qquad
  \Pi_-(p)=\frac{m+\ii\not p}{2m}.
\]
They satisfy \(\Pi_\pm^2=\Pi_\pm\), \(\Pi_+\Pi_-=0\), and
\(\Pi_++\Pi_-=1\).  The defining equations for \(u_R^\sigma\) and
\(v_R^\sigma\) say that they span the \(\Pi_+(p_R)\) and \(\Pi_-(p_R)\)
eigenspaces.

The covariance identity
\[
  R(\Lambda)\gamma^\mu R(\Lambda)^{-1}
  =
  \Lambda^\mu{}_\nu\gamma^\nu
\]
implies
\[
  \not p\,R(L(p)^{-1})=R(L(p)^{-1})\not p_R ,
\]
with \(p=L(p)p_R\).  Applying this identity to \(u_R^\sigma\) and
\(v_R^\sigma\) proves the two momentum-space Dirac equations.

The relation
\[
  R(\Lambda)^\dagger=\beta R(\Lambda^{-1})\beta
\]
is equivalent to \(R(\Lambda)^\dagger\beta R(\Lambda)=\beta\), so the
\(\beta\)-pairing is Lorentz invariant.  Multiplication by
\((m/p^0)^{1/2}\) in the delta-normalized spinors therefore changes the rest
norms only by the factor \(m/p^0\), giving the displayed
\(\beta\)-inner-products.

It remains to identify the matrix-valued completeness relations.  At fixed
\(p\), the vectors \(\mathcal U^\sigma(p)\) form a \(\beta\)-orthogonal basis
of the positive eigenspace, while \(\mathcal V^\sigma(p)\) form a
\(\beta\)-orthogonal basis of the negative eigenspace with negative
\(\beta\)-norm. Hence
\[
  \sum_{\sigma=1}^2
  \mathcal U^\sigma(p)\overline{\mathcal U}^{\,\sigma}(p)
  =
  \frac{m}{p^0}\Pi_+(p),
  \qquad
  \sum_{\sigma=1}^2
  \mathcal V^\sigma(p)\overline{\mathcal V}^{\,\sigma}(p)
  =
  -\frac{m}{p^0}\Pi_-(p).
\]
Substituting the projectors gives the two spin sums with bars. Since
\(\beta^2=1\), the identities without bars follow by multiplying on the right
by \(\beta\). This proves all displayed projector and spin-sum formulas.
\end{proof}

For the explicit gamma matrices displayed above, an adapted rest basis is
\[
  \mathcal U_R^1=\frac{1}{\sqrt2}
  \begin{pmatrix}1\\0\\1\\0\end{pmatrix},
  \qquad
  \mathcal U_R^2=\frac{1}{\sqrt2}
  \begin{pmatrix}0\\1\\0\\1\end{pmatrix},
\]
\[
  \mathcal V_R^1=\frac{1}{\sqrt2}
  \begin{pmatrix}1\\0\\-1\\0\end{pmatrix},
  \qquad
  \mathcal V_R^2=\frac{1}{\sqrt2}
  \begin{pmatrix}0\\1\\0\\-1\end{pmatrix}.
\]
If
\[
  \gamma_5=-\ii\gamma^0\gamma^1\gamma^2\gamma^3
  =
  \begin{pmatrix}
    I_2&0\\ 0&-I_2
  \end{pmatrix},
\]
then
\[
  \mathcal V_R^\sigma=\gamma_5\mathcal U_R^\sigma .
\]
These projector identities are the algebraic source of the external spinor
factors in spinorial LSZ reduction.

\section{Particle Spin and Field Indices}

The spin \(j\) labels an irreducible unitary representation of the massive
little group on the one-particle Hilbert space. A field index \(A\) labels a
finite-dimensional Lorentz representation used by a local operator. The two
are related by the intertwiner
\[
  u_A{}^\sigma(p)
  =
  \bra{\vac}\widehat\Phi_A(0)|\vec p,\sigma\rangle .
\]
This distinction is part of the basic architecture of relativistic quantum
theory. The same spin-\(j\) particle may be created by different local
operators, and a single covariant field representation may contain several
little-group spins when restricted to \(SU(2)\). LSZ for spinning particles
uses the pole of the relevant two-point function together with the
intertwiner \(u_A{}^\sigma(p)\) to convert field indices into external
particle spin labels.

The next chapter constructs spinor field operators and their functional
integral calculus. The representation-theoretic data established here will
then become the spinor completeness relations, propagator numerators, and
external-line factors used in perturbative calculations.
