\chapter{Lorentzian Green Functions and Analytic Continuation}
\label{ch:lorentzian-green-functions}
\ConventionsMixed


\section{Time-Ordered Boundary Values}

\begin{definition}[Lorentzian scalar vacuum-sector datum]
\label{def:lorentzian-scalar-vacuum-sector-datum}
Let \(x^\mu=(x^0,\vec x)\in\mathbb R^{1,D-1}\) and use the mostly-plus
metric
\[
  \eta_{\mu\nu}=\operatorname{diag}(-,+,\ldots,+),
  \qquad
  k^2=\eta_{\mu\nu}k^\mu k^\nu.
\]
Write \(M=\mathbb R^{1,D-1}\).  The framework in this chapter is a scalar
Wightman field presentation in a vacuum sector: a Hilbert space \(\Hilb\), a
vacuum vector \(\vac\), a strongly continuous unitary representation of
translations with self-adjoint generators \(P^\mu\) whose joint spectrum lies
in the closed forward cone
\[
  \overline V_+=\{p\in M:p^0\ge0,\ p^2\le0\},
\]
and a scalar
operator-valued distribution \(\widehat\phi\) defined on a common invariant
dense domain.  Point-field notation denotes the kernel notation for smeared
operator-valued distributions.

For \(r\ge1\), \(\mathcal S(M^r)\) denotes the Schwartz space on \(M^r\), and
\(\mathcal S'(M^r)\) its continuous dual.  This chapter works in the tempered
distribution category:
\[
  W_r,\ G_r\in\mathcal S'(M^r).
\]
Every compactly supported smearing used to define a local field matrix element
is also a Schwartz test function, so this convention is compatible with the
local operator-valued-distribution notation.  After translation invariance is
used, analytic boundary values are paired with test functions in the relative
variables.  Thus a holomorphic function \(F\) on a tube over \(M^{r-1}\) has
boundary value \(T\in\mathcal S'(M^{r-1})\) when, for every
\(f\in\mathcal S(M^{r-1})\),
\[
  \lim_{y\to0}
  \int_{M^{r-1}}\dd^{D(r-1)}\xi\,
  F(\xi-\ii y)f(\xi)
  =
  \langle T,f\rangle ,
\]
with \(y\) approaching the real edge inside the specified tube cone.
\end{definition}

\begin{definition}[Wightman and time-ordered distributions]
\label{def:wightman-time-ordered-distributions}
The Wightman \(n\)-point distribution is
\[
  W_n(x_1,\ldots,x_n)
  =
  \langle \vac|
  \widehat\phi(x_1)\cdots\widehat\phi(x_n)
  |\vac\rangle
  \in \mathcal S'(M^n).
\]
For a permutation \(\pi\in S_n\), set
\[
  M_\pi^n
  =
  \{(x_1,\ldots,x_n):x_{\pi(1)}^0>\cdots>x_{\pi(n)}^0\}.
\]
A vacuum time-ordered Green function for this field is a distribution
\[
  G_n\in\mathcal S'(M^n)
\]
whose restriction to \(M_\pi^n\) is
\[
  G_n(x_1,\ldots,x_n)\big|_{M_\pi^n}
  =
  \langle \vac|
  \widehat\phi(x_{\pi(1)})\cdots
  \widehat\phi(x_{\pi(n)})
  |\vac\rangle .
\]
Equivalently, on \(M_\pi^n\) away from coincident insertion points,
\[
  G_n(x_1,\ldots,x_n)
  =
  \langle \vac|
  \widehat\phi(x_{\pi(1)})\cdots
  \widehat\phi(x_{\pi(n)})
  |\vac\rangle .
\]
Local commutativity makes the restrictions compatible across spacelike
equal-time boundaries.  A choice of \(G_n\) is therefore an extension across
partial diagonals \(x_{a_1}=\cdots=x_{a_r}\).  The possible differences
between two such extensions are local distributions supported on these
diagonals.  In perturbation theory this extension datum is fixed together
with the renormalization prescription for local products.  The compact
operator notation
\[
  G_n(x_1,\ldots,x_n)
  =
  \langle \vac|
  \operatorname T\{
    \widehat\phi(x_1)\cdots\widehat\phi(x_n)
  \}
  |\vac\rangle
\]
refers to this specified distribution.
\end{definition}

\begin{proposition}[Compatibility of time-ordered Wightman pieces]
\label{prop:time-ordered-wightman-piece-compatibility}
Let \(G_n^\circ\) denote the piecewise distribution on the complement of the
partial diagonals \(x_a=x_b\), obtained by assigning the Wightman ordering
corresponding to each strict time-ordering chamber \(M_\pi^n\).  If
\(\widehat\phi\) obeys scalar microcausality,
\[
  [\widehat\phi(f),\widehat\phi(g)]=0
\]
whenever \(\operatorname{supp}f\) and \(\operatorname{supp}g\) are spacelike
separated, then the definitions on adjacent chambers agree across a common
spacelike boundary.  Thus \(G_n^\circ\) is a well-defined distribution on
\[
  M^n\setminus \bigcup_{a<b}\{x_a=x_b\}.
\]
\end{proposition}

\begin{proof}
Let \(U\subset M^n\setminus\bigcup_{a<b}\{x_a=x_b\}\) be a sufficiently small
product coordinate neighborhood.  The equal-time hyperplanes cut \(U\) into
finitely many strict time-ordering chambers.  If two such chambers share a
codimension-one wall in \(U\), then only two labels, say \(a,b\), have been
exchanged; on the wall \(x_a^0=x_b^0\), while \(x_a\neq x_b\) because
partial diagonals have been removed.  After shrinking \(U\), the projections
of \(U\) to the two \(M\)-factors keep the supports of the corresponding
localized test functions spacelike separated.  For such test functions,
microcausality gives
\[
  \widehat\phi(f_a)\widehat\phi(f_b)
  =
  \widehat\phi(f_b)\widehat\phi(f_a)
\]
on the common invariant domain of local-field products.  All remaining field
operators in the smeared product are unchanged, so inserting this equality at
the adjacent positions in the Wightman matrix element shows that the two
chamber formulae have the same restriction to \(U\).

Higher-codimension walls, where several time coordinates coincide but no two
points coincide in spacetime, are handled by the same local argument.  The
permutations labeling the chambers meeting such a wall form a parabolic
subgroup of the symmetric group, generated by adjacent transpositions of
labels whose equal-time points remain spacelike separated in \(U\).  Since
each generator leaves the smeared vacuum matrix element unchanged, any two
chamber prescriptions meeting \(U\) agree on their common restriction.  Thus
the Wightman pieces form a compatible family on the chamber cover of the
off-diagonal open set.  Distributions form a sheaf: a compatible family of
local distributions has a unique glued distribution.  The glued object is
the claimed \(G_n^\circ\).
\end{proof}

\begin{proposition}[Contact-term freedom of the time-ordered extension]
\label{prop:time-ordered-extension-contact-term-freedom}
Let \(G_n\) and \(G_n'\) be two tempered extensions of the same off-diagonal
time-ordered distribution \(G_n^\circ\) to \(M^n\).  Then
\[
  G_n-G_n'
\]
is supported on the union of partial diagonals
\[
  \bigcup_{\mathcal P}
  \left\{x_a=x_b\ \text{whenever \(a,b\) lie in the same block of }
  \mathcal P\right\},
\]
where \(\mathcal P\) ranges over partitions of \(\{1,\ldots,n\}\) with at
least one block of size \(\ge2\).  Locally near a fixed diagonal, this
difference is a finite sum of derivatives of delta distributions in the
normal coordinates, with distributional coefficients along the diagonal.  In a
translation-invariant local renormalization prescription these coefficients
are local polynomials in derivatives acting on delta functions, with finite
coefficients fixed by the chosen extension convention.
\end{proposition}

\begin{proof}
The two extensions agree by hypothesis on the open set
\[
  M^n\setminus \bigcup_{a<b}\{x_a=x_b\}.
\]
Let \(T=G_n-G_n'\).  If a test function is supported in this open set, both
extensions evaluate it by \(G_n^\circ\), hence \(T\) vanishes on such test
functions.  Therefore
\[
  \operatorname{supp}T
  \subset
  \bigcup_{a<b}\{x_a=x_b\}.
\]
Refining this closed cover by the coincidence pattern of the labels gives
the displayed union over partitions \(\mathcal P\): the block structure
records exactly which variables are forced to coincide.

For the local normal form, fix one such stratum and choose coordinates
\((u,v)\), where \(u\) are coordinates along the diagonal and \(v\) are
normal relative coordinates.  A distribution supported on \(v=0\) has finite
transverse order on every compact coordinate chart.  Equivalently, there are
an integer \(N\) and distributions \(T_\alpha\) along the diagonal such that
locally
\[
  \sum_{|\alpha|\le N} T_\alpha(u)\,\partial_v^\alpha\delta(v).
\]
The finite order follows from the seminorm estimate defining continuity of a
distribution on compactly supported tests; Taylor expansion in the normal
variables reduces the action of \(T\) to finitely many normal derivatives at
\(v=0\).  Translation invariance removes dependence on the common
center-of-mass coordinate in flat space.  Locality of the renormalization
prescription further restricts the coefficient distributions to derivatives
of delta functions in relative diagonal variables, with numerical
coupling-dependent coefficients.  These supported delta-derivative
contributions are precisely the contact terms.
\end{proof}

\paragraph{Free scalar Feynman boundary value.}
For the free scalar field of mass \(m>0\), the two-point time-ordered Green
function has Fourier representation
\[
  \Delta_F(x;m^2)
  =
  -\ii
  \int \frac{\dd^D k}{(2\pi)^D}\,
  \frac{\ee^{\ii k\cdot x}}{k^2+m^2-\ii\epsilon},
  \qquad
  \epsilon>0.
\]
At fixed \(\vec k\), with
\[
  \omega_{\vec k}=\sqrt{\vec k^{\,2}+m^2},
\]
the poles in the complex \(k^0\)-plane are
\[
  k^0=-\omega_{\vec k}+\ii0,
  \qquad
  k^0=+\omega_{\vec k}-\ii0 .
\]
This is the Feynman boundary value of the meromorphic function of \(k^0\).
The positive-frequency Wightman two-point function is
\[
  \Delta_+(x;m^2)
  =
  \int\frac{\dd^{D-1}\vec k}{(2\pi)^{D-1}2\omega_{\vec k}}\,
  \ee^{-\ii\omega_{\vec k}x^0+\ii\vec k\cdot\vec x}.
\]
The time-ordered distribution is
\(\theta(x^0)\Delta_+(x)+\theta(-x^0)\Delta_+(-x)\), with the value on
\(x^0=0\) fixed as a distributional extension.  For \(x^0>0\), close the
\(k^0\)-contour of
\[
  -\ii\int\frac{\dd k^0}{2\pi}\,
  \frac{\ee^{-\ii k^0x^0}}{-(k^0)^2+\omega_{\vec k}^2-\ii0}
\]
in the lower half-plane; the pole at \(+\omega_{\vec k}-\ii0\) gives
\(\ee^{-\ii\omega_{\vec k}x^0}/(2\omega_{\vec k})\).  For \(x^0<0\), close
the contour in the upper half-plane; the pole at
\(-\omega_{\vec k}+\ii0\) gives
\(\ee^{+\ii\omega_{\vec k}x^0}/(2\omega_{\vec k})\).  Combining the two time
regions yields the displayed Fourier representation.  The pole placement is
therefore not an independent rule: it is the boundary-value encoding of time
ordering and positive energy.

\section{Tube Analyticity From the Spectrum Condition}

\begin{proposition}[Forward-tube analyticity from the spectrum condition]
\label{prop:forward-tube-analyticity-spectrum-condition}
The \(i\epsilon\) prescription is the two-point perturbative expression of a
more general analytic fact.  Let
\[
  W_n(x_1,\ldots,x_n)
  =
  \langle \vac|
  \widehat\phi(x_1)\cdots\widehat\phi(x_n)
  |\vac\rangle
\]
be a Wightman distribution.  By translation invariance write it in the
relative variables
\[
  \xi_j=x_j-x_{j+1},
  \qquad j=1,\ldots,n-1.
\]
Let
\[
  w_n(\xi_1,\ldots,\xi_{n-1})\in\mathcal S'(M^{n-1})
\]
be the corresponding relative-coordinate distribution.  Its Fourier transform
\(\widetilde w_n\) is defined with the convention
\[
  w_n(\xi)
  =
  \int\frac{\dd^{D(n-1)}p}{(2\pi)^{D(n-1)}}\,
  \ee^{\ii\sum_j p_j\cdot\xi_j}\widetilde w_n(p)
\]
in distribution notation.  Equivalently, for \(f\in\mathcal S(M^{n-1})\),
\[
  \langle w_n,f\rangle
  =
  \left\langle \widetilde w_n,\check f\right\rangle,
  \qquad
  \check f(p)
  =
  \int\frac{\dd^{D(n-1)}\xi}{(2\pi)^{D(n-1)}}\,
  \ee^{\ii\sum_jp_j\cdot\xi_j}f(\xi).
\]
The spectrum condition gives the support inclusion
\[
  \operatorname{supp}\widetilde w_n
  \subset
  \overline V_+^{\,n-1}.
\]
For \(\eta=(\eta_1,\ldots,\eta_{n-1})\in V_+^{n-1}\), the function
\(\exp(\sum_j p_j\cdot\eta_j)\) decreases exponentially on every closed
subcone of \(\overline V_+^{\,n-1}\) which contains the support of
\(\widetilde w_n\).  Multiplication of \(\widetilde w_n\) by this function is
therefore a well-defined tempered distribution after inserting any smooth
cutoff equal to one on a conic neighborhood of the support; the result is
independent of the cutoff.  The inverse Fourier transform is a holomorphic
function of
\begin{equation}
  z_j=\xi_j-\ii\eta_j,
  \qquad
  \eta_j\in V_+,\qquad j=1,\ldots,n-1,
  \label{eq:wightman-forward-tube}
\end{equation}
where \(V_+\) is the open future light cone.  In distribution notation this
Fourier-Laplace transform is written as
\[
  \mathcal W_n(z)
  =
  \frac{1}{(2\pi)^{D(n-1)}}
  \left\langle
    \widetilde w_n(p),
    \exp\!\left\{\ii\sum_{j=1}^{n-1}p_j\cdot z_j\right\}
  \right\rangle .
\]
Indeed, for
\(p\in\overline V_+\) and \(\eta\in V_+\), the mostly-plus inner product
satisfies
\[
  p\cdot\eta\le0,
\]
so the factor
\[
  \exp\{\ii p\cdot(\xi-\ii\eta)\}
  =
  \exp\{\ii p\cdot\xi+p\cdot\eta\}
\]
is damped rather than enlarged.
The boundary value is \(w_n\) in \(\mathcal S'(M^{n-1})\):
\[
  \lim_{\eta\to0,\ \eta_j\in V_+}
  \int_{M^{n-1}}\dd^{D(n-1)}\xi\,
  \mathcal W_n(\xi-\ii\eta)f(\xi)
  =
  \langle w_n,f\rangle,
  \qquad f\in\mathcal S(M^{n-1}).
\]
\end{proposition}

\begin{proof}
Insert a complete spectral resolution of the translation generators between
successive fields.  In relative coordinates the Fourier variables are the
successive momentum transfers carried by these intermediate states.  The
joint spectral condition therefore implies
\(\operatorname{supp}\widetilde w_n\subset\overline V_+^{n-1}\).  If
\(\eta_j\in V_+\) and \(p_j\in\overline V_+\), the mostly-plus metric gives
\(p_j\cdot\eta_j\le0\), with strict negativity away from the boundary unless
one vector is null and parallel to the other.  Hence the factor
\(\exp(\sum_jp_j\cdot\eta_j)\) supplies exponential damping on every closed
subcone inside the tube direction.

For fixed \(\eta\in V_+^{n-1}\), multiplication of the tempered distribution
\(\widetilde w_n\) by the smooth function
\(\exp(\sum_jp_j\cdot\eta_j)\) is defined by first inserting a conic cutoff
equal to one near the support; two such cutoffs differ away from the support
and therefore give the same product.  Differentiation with respect to
\(z_j\) brings down polynomial factors in \(p_j\), which remain valid
multipliers of tempered distributions after the same exponential damping.
Thus the Fourier-Laplace transform is holomorphic in the forward tube.
Letting \(\eta\downarrow0\) inside the tube removes the damping multiplier and
recovers the original Fourier transform in the topology of tempered
distributions, which is exactly the stated boundary value.
\end{proof}

Different operator orderings give holomorphic functions in corresponding
permuted tubes.  Local commutativity identifies their boundary values on real
regions where the exchanged neighboring points are spacelike separated.

\begin{theorem}[Distributional edge-of-the-wedge statement used here]
\label{thm:distributional-edge-of-the-wedge}
Let \(U\subset\mathbb R^N\) be open, and let \(C_+\) and \(C_-\) be open
convex cones in \(\mathbb R^N\) whose convex hull is all of
\(\mathbb R^N\).  Suppose \(F_\pm\) are holomorphic on the tubes
\(U+\ii C_\pm\).  Assume the following polynomial boundary growth: for every
compact \(K\subset U\) and every closed cone \(C'_\pm\) whose nonzero vectors
lie in \(C_\pm\), there are constants \(A,N<\infty\) such that
\[
  |F_\pm(x+\ii y)|
  \le
  A(1+|x|)^N |y|^{-N},
  \qquad
  x\in K,\quad y\in C'_\pm,\quad 0<|y|<1.
\]
These estimates define distributional boundary values
\[
  b_\pm\in\mathcal D'(U).
\]
If \(b_+=b_-\) on a nonempty open real subset \(O\subset U\), then there is a
connected complex neighborhood \(\mathcal N\) of \(O\) and a holomorphic
function \(F\) on \(\mathcal N\) whose restrictions agree with \(F_+\) and
\(F_-\) wherever the original tubes meet \(\mathcal N\).
\end{theorem}

\begin{proof}
This is analytic infrastructure from several complex variables and
distribution theory; the proof is included here to make clear exactly which
part is analytic and which part is supplied by Wightman locality.  The
convex-hull hypothesis is part of the full-neighborhood form of the theorem.
In the more general wedge theorem the continuation lies in the wedge with
imaginary cone
\(\operatorname{conv}(C_+\cup C_-)\); a genuine complex neighborhood of the
real edge is obtained when this cone contains all imaginary directions.  The
permuted Wightman tubes used at a Jost configuration supply precisely this
two-sided normal data after restricting to the relevant spacelike edge.

First, the polynomial tube bounds put each normal translate
\[
  F_{\pm,\epsilon v}:x\longmapsto F_\pm(x+\ii\epsilon v),
  \qquad v\in C_\pm,
\]
in a fixed finite-order distribution class on every \(K\Subset U\).  The
point is not the pointwise bound itself, which may diverge as
\(\epsilon\downarrow0\), but the combination of that bound with Cauchy's
formula.  If \(m\) is larger than the normal growth order, Cauchy's integral
formula in a polydisc of radius comparable to \(\epsilon\) expresses
\(F_{\pm,\epsilon v}\) paired with
\(\varphi\in C_c^\infty(K)\) as a finite sum of pairings of
\(F_\pm(\cdot+\ii\epsilon'v')\) with derivatives of \(\varphi\), where
\(\epsilon'\) remains comparable to \(\epsilon\) and \(v'\) remains in a
proper subcone.  Integrating by parts transfers the singular normal powers to
finitely many derivatives of \(\varphi\).  Thus, for some integer \(M_K\),
\[
  \bigl|\langle F_{\pm,\epsilon v},\varphi\rangle\bigr|
  \le
  C_{K,v}
  \sum_{|\alpha|\le M_K}
  \sup_{x\in K}|\partial^\alpha\varphi(x)|
  \qquad(0<\epsilon<\epsilon_0).
\]
Weak compactness in this finite-order dual gives distributional accumulation
points.  Holomorphy makes the accumulation point unique: if
\(v_0,v_1\in C_\pm\), integrate \(\bar\partial F_\pm=0\) over the truncated
two-real-dimensional surface
\(\{x+\ii\epsilon((1-r)v_0+rv_1):0<\epsilon<\epsilon_0,\ 0<r<1\}\), paired
with a test density in \(x\).  The side contributions vanish as the truncation
shrinks, and the two normal boundary terms are the two candidate limits.
Hence the limit is independent of the approach direction and defines the
boundary distribution \(b_\pm\in\mathcal D'(U)\).

The gluing step is local.  Choose a ball \(B\Subset O\).  In one complex
variable, take a test \((1,1)\)-form supported in a small disc crossing the
real edge and define a distribution by using \(F_+\) in the upper wedge and
\(F_-\) in the lower wedge.  Pairing its \(\bar\partial\) with the test form is
the limit of the contour integral over a rectangle whose horizontal sides lie
at heights \(\pm\epsilon\).  The holomorphic interiors contribute zero.  The
two horizontal boundary terms converge to
\(\langle b_+,\varphi\rangle-\langle b_-,\varphi\rangle\), with the orientation
of the lower side supplying the minus sign.  Since \(b_+=b_-\) on \(B\), the
glued distribution has \(\bar\partial=0\).  The Weyl lemma for the
\(\bar\partial\)-operator, equivalently Morera's theorem in distribution form,
therefore represents it by a holomorphic function across the edge.

For \(N>1\), apply this one-variable argument to complex lines whose real
directions are tangent to \(B\) and whose imaginary directions lie in
\(\operatorname{conv}(C_+\cup C_-)\).  The same boundary-jump cancellation
gives holomorphic continuation on each line.  Separate holomorphy on the
resulting family of line slices and Hartogs' theorem give joint holomorphy on
the local wedge generated by the convex hull.  Bochner's tube theorem then
identifies the envelope of holomorphy of this local tube with the tube over
that convex hull.  Under the hypothesis
\(\operatorname{conv}(C_+\cup C_-)=\mathbb R^N\), the envelope contains a
complex neighborhood of \(B\).  Gluing the local holomorphic extensions over
overlapping balls in \(O\) gives the connected neighborhood \(\mathcal N\).

The pure analytic inputs in this last paragraph are Hartogs' theorem, the
\(\bar\partial\) Weyl lemma, and Bochner's tube theorem.  The QFT-specific
burden in this chapter is the verification of the hypotheses that enter the
analytic theorem: tube growth from the spectrum condition and equality of
boundary distributions from local commutativity.
\end{proof}

\paragraph{Verification of the real-edge equality in the Wightman case.}
Fix two adjacent field labels \(a,b\) and let \(O_{ab}\subset M^n\) be an open
real set on which \(x_a-x_b\) is spacelike and no partial diagonal
\(x_i=x_j\) is met.  Let \(W_{\cdots ab\cdots}\) and
\(W_{\cdots ba\cdots}\) denote the two Wightman distributions that differ
only by interchanging the adjacent fields \(\widehat\phi(x_a)\) and
\(\widehat\phi(x_b)\).  If \(f\in C_c^\infty(O_{ab})\), a partition of unity
subordinate to sufficiently small product neighborhoods writes the smeared
matrix element as a finite sum in which the \(a\)- and \(b\)-smearings have
spacelike separated supports.  Scalar microcausality gives
\[
  \widehat\phi(f_a)\widehat\phi(f_b)
  =
  \widehat\phi(f_b)\widehat\phi(f_a)
\]
on the common invariant domain of the field products appearing in the
vacuum matrix element.  All other smeared fields are unchanged, hence
\[
  \langle W_{\cdots ab\cdots}-W_{\cdots ba\cdots},f\rangle=0,
  \qquad f\in C_c^\infty(O_{ab}).
\]
The boundary values of the two corresponding tube functions therefore agree
as distributions on \(O_{ab}\).  For fields with a \(\mathbb Z_2\)-grading,
the same argument gives the Koszul sign determined by moving the two
homogeneous fields past one another.  This is the real-edge input used by the
distributional edge-of-the-wedge theorem at Jost configurations.

The theorem therefore supplies the common holomorphic continuation of the
permuted tube functions across those spacelike real regions.  Time-ordered
Green functions are boundary values obtained by approaching real Lorentzian
points from imaginary time orderings compatible with the chosen operator
order.  For two points, this general construction is exactly the pole
displacement
\[
  \frac1{k^2+m^2-\ii\epsilon}
\]
appearing in the Feynman propagator.

\section{Boundary-Value Dictionary}

\begin{definition}[Ordered Lorentzian and Euclidean boundary values]
\label{def:ordered-lorentzian-euclidean-boundary-values}
The preceding tube statement gives a precise dictionary among Wightman,
time-ordered, and Euclidean functions.  Fix an ordering \(\pi\) and real
Lorentzian points with
\[
  x_{\pi(1)}^0>x_{\pi(2)}^0>\cdots>x_{\pi(n)}^0 .
\]
Choose positive numbers
\[
  \epsilon_1>\epsilon_2>\cdots>\epsilon_n>0
\]
and set
\[
  z_{\pi(j)}=x_{\pi(j)}-\ii\epsilon_j e_0,
  \qquad e_0=(1,\vec0).
\]
Then
\[
  z_{\pi(j)}-z_{\pi(j+1)}
  =
  x_{\pi(j)}-x_{\pi(j+1)}
  -
  \ii(\epsilon_j-\epsilon_{j+1})e_0
\]
lies in the forward tube for the \(j\)-th relative variable.  The
time-ordered boundary value on the region \(M_\pi^n\) is
\[
  G_n(x_1,\ldots,x_n)\big|_{M_\pi^n}
  =
  \lim_{\epsilon_1>\cdots>\epsilon_n\downarrow0}
  \mathcal W_{n,\pi}(z_{\pi(1)},\ldots,z_{\pi(n)}),
\]
where \(\mathcal W_{n,\pi}\) denotes the holomorphic continuation of the
Wightman ordering displayed by \(\pi\).  The limit is a distributional
boundary value: it is taken after pairing with a Schwartz test function in
the relative variables.

Euclidean Schwinger functions arise from the same holomorphic functions by
restriction to imaginary time.  For Euclidean points
\[
  x_{E,a}=(\tau_a,\vec x_a),
\]
with
\[
  \tau_{\pi(1)}>\tau_{\pi(2)}>\cdots>\tau_{\pi(n)},
\]
define
\[
  z_{\pi(j)}^0=-\ii\tau_{\pi(j)},
  \qquad
  \vec z_{\pi(j)}=\vec x_{\pi(j)} .
\]
Then
\[
\begin{aligned}
  &S_n(x_{E,1},\ldots,x_{E,n})
    \big|_{\tau_{\pi(1)}>\cdots>\tau_{\pi(n)}}  \\
  &\qquad =
  \mathcal W_{n,\pi}(z_{\pi(1)},\ldots,z_{\pi(n)}).
\end{aligned}
\]
\end{definition}

\paragraph{Source sign under Wick rotation.}
The same dictionary fixes the sign in source functionals.  Let
\[
  Z_L^{(+)}[J]
  =
  \bra{\vac}
  \operatorname T\exp\left(
    \ii\int_{\mathbb R^{1,D-1}}J(x)\widehat\phi(x)\,\dd^Dx
  \right)
  \ket{\vac}
\]
denote the Lorentzian source convention in which functional derivatives
produce factors of \(\ii\).  Its opposite-source convention is
\[
  Z_L^{(-)}[J]=Z_L^{(+)}[-J].
\]
To compare with a Euclidean source one must specify the contour, because a
real Euclidean test function and a real Lorentzian test function live on
different real slices.  Let \(C_-\) be the imaginary-time contour
\[
  z^0=-\ii\tau,\qquad \dd z^0=-\ii\,\dd\tau,
\]
with the orientation inherited from increasing \(\tau\), and let \(J_C\) be a
source density defined on this contour.  Term by term in the source expansion,
or in a regulated theory where the contour-ordered source exponential is an
operator-valued analytic function of the source, the source part of the
exponent is
\[
  \ii\int_{C_-}\dd z^0\,\dd^d\vec x\,
    J_C(z^0,\vec x)\widehat\phi(z^0,\vec x)
  =
  \int_{\mathbb R}\dd\tau\,\dd^d\vec x\,
    J_C(-\ii\tau,\vec x)
    \widehat\phi_E(\tau,\vec x).
\]
Thus the Euclidean convention
\[
  Z_E[J_E]
  =
  \bra{\vac}
  T_E\exp\left(
    \int J_E(\tau,\vec x)
      \widehat\phi_E(\tau,\vec x)\,\dd\tau\,\dd^d\vec x
  \right)
  \ket{\vac}
\]
is the \(C_-\) boundary value of \(Z_L^{(+)}\) with
\[
  J_E(\tau,\vec x)=J_C(-\ii\tau,\vec x).
\]
No additional minus sign is produced in this convention; the factor
\(\ii\) in the Lorentzian source and the Jacobian
\(-\ii\) of the lower-half-plane Wick contour multiply to \(+1\).  If one
instead rotates through the opposite contour \(C_+\), \(z^0=+\ii\tau\), then
\(\dd z^0=\ii\,\dd\tau\) and the same Lorentzian convention gives
\[
  \ii\int_{C_+}\dd z^0\,J_C\widehat\phi
  =
  -\int\dd\tau\,J_C(\ii\tau)\widehat\phi_E .
\]
In that contour convention the Euclidean functional with
\(+\int J_E\phi_E\) is represented by \(Z_L^{(+)}[-J_C]\), equivalently by
\(Z_L^{(-)}[J_C]\), after setting \(J_E(\tau,\vec x)=-J_C(\ii\tau,\vec x)\).
The two formulas describe different contour/source-sign conventions, not two
different Green functions.
Expand the contour-ordered exponential in powers of the source.  Each
Lorentzian source insertion carries the factor \(\ii\,\dd z^0\).  On the
lower imaginary-time contour \(z^0=-\ii\tau\), the one-dimensional volume form
is \(\dd z^0=-\ii\,\dd\tau\), and therefore
\(\ii\,\dd z^0=+\dd\tau\).  The source term becomes \(+\int J_E\phi_E\) after
the identification \(J_E(\tau,\vec x)=J_C(-\ii\tau,\vec x)\).  On the upper
contour \(z^0=+\ii\tau\), one instead has
\(\ii\,\dd z^0=-\dd\tau\), so the same Lorentzian source convention produces
the opposite Euclidean sign unless the source is replaced by its negative.
The calculation is termwise in the regulated source expansion, hence fixes
all Euclidean \(n\)-point signs.

Conversely, an OS-admissible Schwinger hierarchy reconstructs Wightman
boundary values by the positive-time quotient and analytic-continuation
theorem of Chapter~\ref{ch:volume-iv-osterwalder-schrader-reconstruction}.
Thus the boundary-value dictionary has two controlled entrances: spectral
Wightman data with tube analyticity, and Euclidean data satisfying the OS
positivity, covariance, regularity, and growth hypotheses.

\paragraph{Two-point Feynman prescription from ordered tubes.}
For \(n=2\), the dictionary gives the Feynman prescription.  If
\(x^0>0\), the ordered boundary value is obtained with the positive-energy
pole below the real \(k^0\)-axis and the negative-energy pole above it:
\[
  k^0=+\omega_{\vec k}-\ii0,
  \qquad
  k^0=-\omega_{\vec k}+\ii0.
\]
If \(x^0<0\), the contour closure selects the opposite residue in the same
single distribution \(\Delta_F\).  This two-point pole rule is the momentum
space encoding of the ordered tube boundary value above.
The ordered tube prescription for \(x^0>0\) approaches the first field from a
larger negative imaginary time than the second.  In Fourier variables this is
the same damping choice as in the free two-point boundary-value calculation
above, so the positive-energy pole is passed below the real axis.  If
\(x^0<0\), the operator order is reversed and the contour closure selects the
negative-energy pole above the axis.  The two cases are precisely the two
boundary values of the single
meromorphic denominator \(k^2+m^2\), hence the denominator is
\(k^2+m^2-\ii0\).

\section{Lorentzian Feynman Rules for Green Functions}

\begin{definition}[Lorentzian labelled graph rules for Green functions]
\label{def:lorentzian-labelled-graph-rules-green-functions}
Consider the same scalar \(\phi^4\) interaction as in the Euclidean chapter,
with Lorentzian interaction density
\[
  \mathcal L_{\mathrm{int}}
  =
  -\frac{g}{4!}\phi^4 .
\]
The perturbative rules below compute time-ordered Green functions. They use
Lorentzian momenta on every line.

\begin{center}
\begin{tikzpicture}[scale=0.9,>=Stealth]
  \draw[purple!80!black,very thick,->] (-1.25,0)--(1.25,0)
    node[midway,above] {$k$};
  \node at (2.0,0) {$=$};
  \node at (3.35,0)
    {$\displaystyle \frac{-\ii}{k^2+m^2-\ii\epsilon}$};

  \begin{scope}[xshift=7.2cm]
    \fill[blue!80!black] (0,0) circle (2.5pt);
    \foreach \ang/\lab in {135/k_1,45/k_2,-45/k_3,-135/k_4} {
      \draw[purple!80!black,very thick,->]
        ({1.0*cos(\ang)},{1.0*sin(\ang)}) -- (0,0);
      \node at ({1.34*cos(\ang)},{1.34*sin(\ang)}) {$\lab$};
    }
    \node at (1.85,0) {$=$};
    \node[align=left,anchor=west] at (2.25,0)
      {$\displaystyle -\ii g\,(2\pi)^D
        \delta^{(D)}(k_1+k_2+k_3+k_4)$};
  \end{scope}
\end{tikzpicture}
\end{center}

This is the labelled-leg convention matching the Euclidean convention of the
previous chapter. Direct expansion of the interaction density instead assigns
\[
  -\frac{\ii g}{4!}\,(2\pi)^D
  \delta^{(D)}(k_1+k_2+k_3+k_4)
\]
to an unlabelled four-field insertion and recovers the labelled rule by
counting contractions.
In the labelled-leg convention used below, the factor \(4!\) associated with
assigning the four incident half-edges to a quartic vertex has already been
absorbed into the vertex rule \(-\ii g\).  Therefore a connected graph
\(\Gamma\) with labelled external legs carries only the residual factor
\[
  \frac{1}{|\operatorname{Aut}_{\mathrm{ext}}\Gamma|},
\]
where \(\operatorname{Aut}_{\mathrm{ext}}\Gamma\) is the automorphism group of
the graph that fixes every external label and preserves the incidence
relations.  For the tadpole graph below this residual group has order \(2\),
coming from the interchange of the two internal half-edges forming the loop;
this is why the tadpole coefficient is \(-\ii g/2\), not \(-\ii g\).
\end{definition}

\paragraph{Lorentzian--Euclidean graph factor under simultaneous rotation.}
Consider a connected \(\phi^4\) graph contributing to a time-ordered
\(n\)-point Green function with external propagators included.  Suppose the
external momenta lie in the first-sheet Euclidean domain and the internal loop
energy contours can be simultaneously deformed to imaginary axes without
crossing or pinching singularities.  Let \(I\) be the number of internal
propagators, \(V\) the number of quartic vertices, and
\[
  L=I-V+1
\]
the number of loop momenta.  With the labelled conventions of this chapter
and Chapter~\ref{ch:perturbative-green-functions}, the Lorentzian graph
boundary value at \(k_a^0=\ii k_{E,a}^D\) is
\[
  (-1)^n\ii^{\,n+1}
\]
times the corresponding Euclidean connected Green-function graph.  In
particular, a connected two-point graph gives the factor \(-\ii\), while a
connected four-point graph gives the factor \(+\ii\).
The Lorentzian labelled graph contains one factor \(-\ii g\) for each quartic
vertex, one factor
\[
  \frac{-\ii}{p_e^2+m^2-\ii0}
\]
for each propagator edge, and one real loop measure for each independent loop
momentum.  In a connected \(n\)-point Green function with external propagators
included, the total number of propagator edges is \(I+n\), because each
external insertion is attached to the interaction graph by one external
propagator.  Rotating each independent loop energy by
\(\ell^0=\ii\ell_E^D\) gives one Jacobian factor \(\ii\) for each loop.
The denominator of every propagator becomes the corresponding Euclidean
denominator by the assumed contour deformation.

Thus the total phase multiplying the Euclidean denominators and the factor
\(g^V\) is
\[
  (-\ii)^V(-\ii)^{I+n}\ii^L .
\]
The Euclidean graph with the same labelled convention has vertex factor
\((-g)^V\), the same symmetry factor, and no propagator numerator phase.
Dividing by the Euclidean vertex phase gives the ratio
\[
  \frac{(-\ii)^{V+I+n}\ii^{I-V+1}}{(-1)^V}
  =
  (-1)^n\ii^{\,n+1},
\]
where \(L=I-V+1\) has been used.  The result is independent of the number of
vertices and internal lines.  The assumption on contour deformation is what
excludes threshold pinches; without it the graph is defined by boundary value
on the cut rather than by this direct simultaneous rotation.

\section{Loop Rotation and the Tadpole}

\paragraph{Tadpole contour rotation and self-energy matching.}
The one-loop tadpole contribution to the Lorentzian two-point self-energy
insertion is
\[
\begin{tikzpicture}[baseline=-0.45ex,scale=0.75,>=Stealth]
  \draw[purple!80!black,very thick,->] (-1.1,0)--(0,0)
    node[midway,below] {$k$};
  \draw[purple!80!black,very thick,->] (1.1,0)--(0,0)
    node[midway,below] {$k$};
  \fill (0,0) circle (2.2pt);
  \draw[purple!80!black,very thick,->]
    (0,0) .. controls (-0.35,0.9) and (0.35,0.9) .. (0,0)
    node[midway,above] {$\ell$};
\end{tikzpicture}
\quad=\quad
-\frac{\ii g}{2}
\int \frac{\dd^D\ell}{(2\pi)^D}\,
\frac{-\ii}{\ell^2+m^2-\ii\epsilon}.
\]
The factor \(1/2\) is the same tadpole symmetry factor as in Euclidean
signature.  This is a combinatorial statement: the Wick contraction count
depends only on the polynomial \(\phi^4\) and on the residual graph
automorphism, while the signature-dependent factors enter separately through
the labelled Lorentzian vertex \(-\ii g\) and propagator numerator \(-\ii\).
The loop energy \(\ell^0\) is integrated along the real axis with
the Feynman pole displacement. The Euclidean integral is obtained by rotating
the contour to \(\ell^0=\ii \ell_E^D\), with the contour passing between the
two displaced poles.

\begin{center}
\begin{tikzpicture}[scale=0.95,>=Stealth]
  \draw[->,line width=0.65pt] (-3.1,0)--(3.1,0)
    node[right] {$\operatorname{Re}\ell^0$};
  \draw[->,line width=0.65pt] (0,-2.05)--(0,2.05)
    node[above] {$\operatorname{Im}\ell^0$};
  \draw[blue!70!black,very thick,->] (-2.75,0)--(2.75,0);
  \draw[green!50!black,very thick,->] (0,-1.75)--(0,1.75);
  \node[green!50!black,anchor=west,font=\small] at (0.18,1.82)
    {$\ell^0=\ii\ell_E^D$};
  \node[red!80!black] at (-1.35,0.35) {$\times$};
  \node[red!80!black] at (1.35,-0.35) {$\times$};
  \node[red!80!black,anchor=east] at (-1.52,0.92)
    {$-\omega_{\vec\ell}+\ii0$};
  \node[red!80!black,anchor=west] at (1.52,-1.02)
    {$+\omega_{\vec\ell}-\ii0$};
  \draw[orange!85!black,very thick,->]
    (0.38,1.45) .. controls (1.25,1.12) and (1.95,0.65) .. (2.45,0.08);
  \draw[orange!85!black,very thick,->]
    (-0.38,-1.45) .. controls (-1.25,-1.12) and (-1.95,-0.65) .. (-2.45,-0.08);
  \node[orange!85!black,font=\small,anchor=west] at (2.25,1.35)
    {contour rotation};
\end{tikzpicture}
\end{center}

With this orientation,
\[
\begin{gathered}
  \ell^0=\ii\ell_E^D,\qquad
  \dd\ell^0=\ii\,\dd\ell_E^D,\qquad
  \dd^D\ell=\ii\,\dd^D\ell_E,\\
  \ell^2+m^2-\ii\epsilon
  =
  -(\ell^0)^2+\vec\ell^{\,2}+m^2-\ii\epsilon
  \longrightarrow
  \ell_E^2+m^2 .
\end{gathered}
\]
Therefore the measure and propagator numerator together give
\[
  \dd^D\ell\,\frac{-\ii}{\ell^2+m^2-\ii\epsilon}
  \longrightarrow
  \ii\,\dd^D\ell_E\,\frac{-\ii}{\ell_E^2+m^2}
  =
  \dd^D\ell_E\,\frac{1}{\ell_E^2+m^2}.
\]
Including the vertex factor, this gives
\[
\begin{aligned}
  -\frac{\ii g}{2}
  \int \frac{\dd^D\ell}{(2\pi)^D}
  \frac{-\ii}{\ell^2+m^2-\ii\epsilon}
  &=
  \ii\left[
    -\frac{g}{2}
    \int \frac{\dd^D\ell_E}{(2\pi)^D}
    \frac{1}{\ell_E^2+m^2}
  \right].
\end{aligned}
\]
Thus the Lorentzian two-point insertion is \(\ii\Sigma(k)\), where
\(\Sigma(k)\) is the first-sheet analytic continuation of the Euclidean
self-energy \(\Sigma_E(k_E)\) with the sign convention of
Chapter~\ref{ch:perturbative-green-functions}.
More explicitly, in the same regulator and the same local two-point
subtraction convention, the denominator self-energy obeys
\begin{equation}
  \Sigma\bigl(k^0=\ii k_E^D,\vec k=\vec k_E\bigr)
  =
  \Sigma_E(k_E),
  \label{eq:lorentzian-euclidean-self-energy-match}
\end{equation}
where \(\Sigma_E\) is defined by
Eq.~\eqref{eq:euclidean-self-energy}.  The factor \(\ii\) displayed above
belongs to the Lorentzian two-point insertion in the Dyson series; it is not an
extra sign in the scalar function \(\Sigma\) appearing in the denominator of
\(\widetilde G(k)\).
The coefficient \(-\ii g/2\) follows from
Definition~\ref{def:lorentzian-labelled-graph-rules-green-functions}: the
labelled quartic vertex supplies \(-\ii g\), and the loop made from the two
remaining half-edges has residual automorphism group of order \(2\).  The
Feynman pole displacement places the positive-energy pole below and the
negative-energy pole above the real loop-energy axis.  Rotating the real
\(\ell^0\)-cycle to \(\ell^0=\ii\ell_E^D\) crosses no pole in this tadpole
because no external energy flows through the loop.  The Jacobian gives
\(\dd^D\ell=\ii\dd^D\ell_E\), while the propagator numerator gives
\(-\ii\); their product is \(+1\).  The remaining vertex factor is exactly
\(\ii\) times the Euclidean tadpole insertion.  Since the Lorentzian
two-point Green function is written with a global propagator numerator
\(-\ii\), this extra \(\ii\) belongs to the insertion in the Dyson series and
not to the denominator self-energy scalar itself.  This gives
\eqref{eq:lorentzian-euclidean-self-energy-match}.

\begin{proposition}[Loop rotation and pinch obstruction]
\label{prop:loop-rotation-pinch-obstruction}
The contour rotation just used is a special case with no external energy
flowing through the loop.  The same direct rotation is also legitimate for
loop integrals with Euclidean external momenta, where the external energies
lie on the imaginary axis and the Feynman poles can be separated continuously
from the rotating contours.  For general Lorentzian timelike external momenta,
one must instead analyze the full contour deformation in the complex loop
energy variables.  The obstruction is a pinch singularity: poles from different
propagators can trap the integration cycle so that it cannot be deformed to
the Euclidean cycle without crossing singularities.

For example, the Lorentzian continuation of the order-\(g^2\) sunset
self-energy contains the denominator product
\[
  \frac{1}{
  (q_1^2+m^2-\ii0)(q_2^2+m^2-\ii0)
  \bigl((k-q_1-q_2)^2+m^2-\ii0\bigr)} .
\]
At \(\vec k=0\), the Landau equations have a solution with all three internal
lines on shell and carrying positive energy when
\[
  (k^0)^2\ge (3m)^2 .
\]
At and above this three-particle threshold the \(q_1^0,q_2^0\) contours are
pinched.  Thus the Euclidean sunset integral defines a first-sheet analytic
function away from its Landau singularities, and the physical Lorentzian value
above threshold is the specified boundary value on the cut, not the result of
a naive independent rotation of each loop energy.  The discontinuity across
such cuts is computed by the Cutkosky rules, which are the perturbative
expression of the same spectral and unitarity structure developed in
Chapter~\ref{ch:volume-ii-analyticity-crossing-landau}.
\end{proposition}

\begin{proof}
For Euclidean external energies, every external energy entering a propagator
has imaginary value.  For fixed spatial loop momenta, the pole locations in
each loop-energy variable remain a positive distance away from the rotating
contour on compact subsets of the Euclidean external domain.  The Feynman
\(i0\) prescription places the positive- and negative-energy poles on
opposite sides of the original real contour, and the imaginary external
energies prevent a pole from crossing the homotopy from the real contour to
the imaginary contour.  Cauchy's theorem for the regularized meromorphic
integrand identifies the Lorentzian integral with the Euclidean integral on
this first-sheet domain.  Equivalently, insert a rapidly decreasing
regulator, perform the contour deformation with vanishing large arcs, and
remove the regulator inside the common analytic domain.

For real timelike \(k\), the sunset denominators admit the simultaneous
on-shell solution
\[
  q_1^2=q_2^2=(k-q_1-q_2)^2=-m^2,
  \qquad
  q_1^0,q_2^0,k^0-q_1^0-q_2^0>0
\]
when \(k^0\ge3m\) in the rest frame.  At threshold the three internal spatial
momenta may be zero; above threshold they may share the total energy
\(k^0\).  The Landau stationarity condition says that the gradients of the
three on-shell denominators are linearly dependent with nonnegative
coefficients, so the singular hypersurfaces meet the integration cycle in a
trapping configuration.  In the \(q_1^0,q_2^0\) variables, poles
approach from opposite sides as the external energy is continued to the
physical value.  No deformation to the imaginary cycle exists without
crossing a pole.  The Lorentzian value above threshold is therefore a
specified boundary value on the cut of the first-sheet analytic function, not
the result of an independent Euclidean rotation of each loop energy.
\end{proof}

\section{Full Two-Point Function and First Sheet}

\begin{definition}[First-sheet two-point continuation]
\label{def:first-sheet-two-point-continuation}
The perturbative time-ordered two-point function is written as
\[
  G(x,y)
  =
  \int\frac{\dd^D k}{(2\pi)^D}\,
  \ee^{\ii k\cdot(x-y)}
  \widetilde G(k),
\]
where
\[
  \widetilde G(k)
  =
  \frac{-\ii}{k^2+m^2-\ii\epsilon-\Sigma(k)} .
\]
The function \(\Sigma(k)\) is initially computed for Euclidean external
momenta, where loop integrals have Euclidean denominators. Using
\[
  k^0=\ii k_E^D ,
\]
the Euclidean external energy lies on the imaginary \(k^0\)-axis. The
matching identity \eqref{eq:lorentzian-euclidean-self-energy-match} gives
\[
  \widetilde G(\ii k_E^D,\vec k_E)
  =
  -\ii\,\widetilde G_E(k_E),
\]
because \(k^2=k_E^2\) at \(k^0=\ii k_E^D\) with the mostly-plus convention.
The Lorentzian value is the boundary value obtained by analytic continuation of
\(k^0\) from this axis to the prescribed side of the real axis without crossing
the first-sheet cuts.
\end{definition}

\paragraph{Spectral thresholds and first-sheet cuts.}
The spectral representation fixes the location of the first singular
thresholds. Suppose the continuum part \(d\rho_{\mathrm c}\) in the two-point
spectral measure has lowest invariant mass \(M\), meaning
\[
  M^2
  =
  \inf \operatorname{supp} d\rho_{\mathrm c}
\]
in the K\"all\'en--Lehmann variable \(s=\mu^2\).  Under the two-particle
threshold assumption stated in Chapter~\ref{ch:kallen-lehmann}, namely
\(\operatorname{supp}d\rho_{\mathrm c}\subset[4m^2,\infty)\) with the
\(\phi\phi\) continuum saturating the lower endpoint, this notation gives
\[
  M^2=4m^2,\qquad M=2m .
\]
If the first continuum channel has a different threshold, \(M\) denotes that
actual square-root threshold. At fixed spatial momentum \(\vec k\), define
\[
  E_M(\vec k)=\sqrt{\vec k^{\,2}+M^2}.
\]
The continuum denominators have singular support when
\[
  (k^0)^2=\vec k^{\,2}+s,
  \qquad s\ge M^2.
\]
Consequently the first sheet has branch cuts beginning at
\[
  k^0=+E_M(\vec k),
  \qquad
  k^0=-E_M(\vec k).
\]
The Feynman boundary value on the positive-energy cut is reached from the
upper half-plane. Thus the value of \(\Sigma(k)\) at real positive \(k^0\) is
the boundary value approached from \(\operatorname{Im}k^0>0\) on the first
sheet. The negative-energy cut is reached with the conjugate placement.
The reason is direct but important.  The K\"all\'en--Lehmann representation
writes the connected two-point
function as an integral of free denominators over the positive spectral
measure:
\[
  \int_{[M^2,\infty)}\frac{\dd\rho_{\rm c}(s)}
  {k^2+s-\ii0}.
\]
At fixed \(\vec k\), each value of \(s\) has branch points or poles at
\(k^0=\pm\sqrt{\vec k^{\,2}+s}\).  The union of these singular loci over
\(s\ge M^2\) is exactly the pair of cuts beginning at
\(\pm E_M(\vec k)\).  The Feynman prescription puts the positive-energy
denominator below the real axis, so the physical positive-energy boundary
value is obtained by approaching the cut from \(\operatorname{Im}k^0>0\).
The negative-energy side is the conjugate boundary value imposed by the same
real spectral measure.

\begin{center}
\begin{tikzpicture}[scale=0.95,>=Stealth]
  \draw[->,line width=0.65pt] (-3.15,0)--(3.15,0)
    node[right] {$\operatorname{Re}k^0$};
  \draw[->,line width=0.65pt] (0,-2.05)--(0,2.05)
    node[above] {$\operatorname{Im}k^0$};
  \draw[red!80!black,very thick,decorate,
    decoration={zigzag,segment length=4pt,amplitude=1.2pt}]
    (-3.0,0)--(-1.05,0);
  \draw[red!80!black,very thick,decorate,
    decoration={zigzag,segment length=4pt,amplitude=1.2pt}]
    (1.05,0)--(3.0,0);
  \node[below] at (-1.05,-0.05) {$-E_M(\vec k)$};
  \node[below] at (1.05,-0.05) {$+E_M(\vec k)$};
  \draw[orange!85!black,very thick,->]
    (1.65,1.05) .. controls (2.2,0.82) and (2.2,0.28) .. (1.65,0.12);
  \draw[orange!85!black,very thick,->]
    (-1.65,-1.05) .. controls (-2.2,-0.82) and (-2.2,-0.28) .. (-1.65,-0.12);
  \node[orange!85!black,align=center] at (2.25,1.34)
    {positive-energy\\boundary value};
  \node[orange!85!black,align=center] at (-2.25,-1.34)
    {negative-energy\\boundary value};
\end{tikzpicture}
\end{center}

The output of the continuation is a family of Lorentzian time-ordered Green
functions with specified singularity prescriptions. The spectral poles and
cuts identify one-particle and continuum invariant masses seen by the field.
The next construction concerns asymptotic states: it asks for Hilbert-space
limits in which separated wave packets behave as freely propagating stable
particles in the far past or far future. That construction is logically
additional to the analytic continuation of Green functions.
