\chapter{Integrable RG Flows and Perturbed Two-Dimensional CFT}
\label{ch:integrable-rg-flows-perturbed-cft}

\ConventionsMixedCFT

This chapter studies two-dimensional quantum field theories specified near a
conformal fixed point by a relevant scalar local operator and studied away
from the fixed point through factorized scattering, form factors, and
finite-size spectra.  The word flow will always mean a comparison between
renormalized local QFT data at different length scales.  An identification of
such a flow includes a declared local-operator chart, infinite-volume
scattering data, and a finite-volume Hilbert-space construction whose
ground-state energy is computed by the stated TBA system.

\section{Perturbed-CFT Data}

Let \(\mathcal C\) be a Euclidean two-dimensional CFT on the plane.  A scalar
primary operator \(\mathcal O\) has holomorphic and antiholomorphic weights
\((h,\bar h)\) with \(h=\bar h\), and scaling dimension
\[
  \Delta_{\mathcal O}=h+\bar h=2h .
\]
It is relevant if
\[
  y_{\mathcal O}:=2-\Delta_{\mathcal O}>0 .
\]
The number \(y_{\mathcal O}\) is the engineering dimension of the coupling
that multiplies the integrated insertion \(\int d^2x\,\mathcal O(x)\).  This
dimension statement is independent of perturbation theory: under the Euclidean
dilation \(x\mapsto \lambda x\), the measure contributes \(\lambda^2\) and
the primary insertion contributes \(\lambda^{-\Delta_{\mathcal O}}\).

\begin{definition}[Renormalized perturbed-CFT chart]
\label{def:renormalized-perturbed-cft-chart}
A renormalized perturbation chart for \((\mathcal C,\mathcal O)\) consists of
the following data.

\begin{enumerate}
\item A normalization of \(\mathcal O\), for example
\[
  \langle \mathcal O(x)\mathcal O(0)\rangle_{\mathcal C}
  =
  \frac{1}{|x|^{2\Delta_{\mathcal O}}}
  \quad (x\neq0),
\]
or another explicitly stated nonzero constant in the numerator.
\item A local source coordinate \(g\) with mass dimension
\([g]=y_{\mathcal O}\).
\item For each collection \(X\) of separated spectator insertions and each
integer \(n\ge0\), an extension prescription
\[
  \mathcal R_n\!\left[
  \langle X\,\mathcal O(x_1)\cdots\mathcal O(x_n)\rangle_{\mathcal C}
  \right]
\]
from the configuration space where all points are distinct to a distribution
on the full product space.  The extension prescription must state its contact
terms on diagonals; different choices are different renormalized coordinates.
\end{enumerate}

The formal perturbation series in this chart is
\[
  \langle X\rangle_g
  =
  \frac{
  \sum_{n\ge0}\frac{(-g)^n}{n!}
  \int \prod_{j=1}^n d^2x_j\,
  \mathcal R_n[
  \langle X\,\mathcal O(x_1)\cdots\mathcal O(x_n)\rangle_{\mathcal C}]}
  {\sum_{n\ge0}\frac{(-g)^n}{n!}
  \int \prod_{j=1}^n d^2x_j\,
  \mathcal R_n[
  \langle \mathcal O(x_1)\cdots\mathcal O(x_n)\rangle_{\mathcal C}]} .
\]
The expression is a formal or regulated object until the volume, infrared
boundary condition, and convergence or summability statement have been
specified.
\end{definition}

The extension maps in Definition~\ref{def:renormalized-perturbed-cft-chart}
are part of the definition of the perturbed theory.  They are where contact
terms enter a perturbed CFT.  In a distributional formulation the product of
local fields is first defined away from diagonals.  A renormalized
perturbation chooses extensions to the diagonals subject to locality,
covariance, and the declared source normalization.  Thus the coupling \(g\) is
a coordinate in the Lagrangian or source functional of a specified regularized
and renormalized path integral.

\begin{proposition}[First-order scaling of a relevant source]
\label{prop:first-order-relevant-source-scaling}
Let \(g(\mu)\) denote the dimensionful source in the chart above, and let
\[
  \lambda(\mu)=\mu^{-y_{\mathcal O}}g(\mu)
\]
be the corresponding dimensionless coordinate at renormalization scale
\(\mu\).  If contact-term extensions are not changed while comparing two
nearby scales and nonlinear OPE contributions are ignored, then
\[
  \mu\frac{d\lambda}{d\mu}=-y_{\mathcal O}\lambda .
\]
Equivalently, increasing the length scale \(\ell=\log(\mu_0/\mu)\) gives
\[
  \frac{d\lambda}{d\ell}=y_{\mathcal O}\lambda .
\]
\end{proposition}

\begin{proof}
The dimensionful source \(g\) has mass dimension \(y_{\mathcal O}\).  Holding
the source functional fixed while changing only the unit of mass gives
\(\lambda(\mu)=\mu^{-y_{\mathcal O}}g\).  Differentiating this identity at
fixed \(g\) proves the first displayed equation.  The second follows from
\(\ell=-\log\mu+\mathrm{constant}\).  Corrections beyond this statement come
from the diagonal extension of products of \(\mathcal O\), equivalently from
the singular part of the \(\mathcal O\times\mathcal O\) OPE in the chosen
renormalized coordinate.
\end{proof}

If the perturbation generates a massive theory with mass gap \(M>0\), the
dimensionless finite-size variable on a circle of circumference \(R\) is
\[
  r=MR .
\]
Dimensional analysis fixes only the power relating \(M\) and
\(|g|^{1/y_{\mathcal O}}\).  Once the normalization of \(\mathcal O\), the
source coordinate, and the scattering mass convention have been declared, an
integrable model may determine a numerical mass-coupling constant
\[
  M=\kappa_{\mathcal O}\,|g|^{1/y_{\mathcal O}} .
\]
The number \(\kappa_{\mathcal O}\) is part of the matching data after the
coordinates are fixed.

\section{Unitary Minimal Models and \texorpdfstring{\(\phi_{1,3}\)}{phi13}}

The A-series unitary Virasoro minimal model is denoted
\(\mathcal M(m,m+1)\), with \(m\ge3\).  Its central charge is
\[
  c_m=1-\frac{6}{m(m+1)} ,
\]
and its chiral Kac weights are
\[
  h_{r,s}^{(m)}
  =
  \frac{\big((m+1)r-ms\big)^2-1}{4m(m+1)},
  \qquad
  1\le r\le m-1,\quad 1\le s\le m ,
\]
with the identification \((r,s)\sim(m-r,m+1-s)\).  The diagonal full CFT has
scalar primaries when the same chiral module is paired with its
antiholomorphic copy.

\begin{proposition}[Elementary data of the \(\phi_{1,3}\) perturbation]
\label{prop:phi13-minimal-model-data}
In \(\mathcal M(m,m+1)\), the diagonal scalar primary
\(\phi_{1,3}\) has
\[
  h_{1,3}^{(m)}=\frac{m-1}{m+1},
  \qquad
  \Delta_{1,3}^{(m)}=\frac{2(m-1)}{m+1},
  \qquad
  y_{1,3}^{(m)}=\frac{4}{m+1}.
\]
Moreover
\[
  c_m-c_{m-1}=\frac{12}{m(m^2-1)}
  \qquad (m\ge4).
\]
\end{proposition}

\begin{proof}
For \((r,s)=(1,3)\),
\[
  (m+1)r-ms=m+1-3m=1-2m .
\]
Therefore
\[
  h_{1,3}^{(m)}
  =
  \frac{(2m-1)^2-1}{4m(m+1)}
  =
  \frac{4m(m-1)}{4m(m+1)}
  =
  \frac{m-1}{m+1}.
\]
The full diagonal scalar has scaling dimension \(2h_{1,3}^{(m)}\), giving the
displayed \(\Delta_{1,3}^{(m)}\).  Subtracting from \(2\) gives
\[
  y_{1,3}^{(m)}
  =
  2-\frac{2(m-1)}{m+1}
  =
  \frac{4}{m+1}.
\]
Finally,
\[
  c_m-c_{m-1}
  =
  \left(1-\frac{6}{m(m+1)}\right)
  -
  \left(1-\frac{6}{(m-1)m}\right)
  =
  \frac{6}{m}\left(\frac1{m-1}-\frac1{m+1}\right),
\]
which is \(12/[m(m^2-1)]\).
\end{proof}

The standard integrable identification asserts that, for the massless branch
of the \(\phi_{1,3}\) perturbation and a definite sign convention for the
source \(g\),
\[
  \mathcal M(m,m+1)
  +g\int d^2x\,\phi_{1,3}(x)
  \quad\longrightarrow\quad
  \mathcal M(m-1,m)
  \qquad (m\ge4).
\]
In this monograph the theorem boundary is kept explicit.  The exact
arithmetic in Proposition~\ref{prop:phi13-minimal-model-data} proves only
that the proposed infrared central charge is smaller by the displayed
positive amount and that the perturbing operator is relevant.  A theorem-level
identification of the flow would require a construction of the perturbed
local QFT, a proof of its infrared limit, and a matching of its local operator
algebra to that of \(\mathcal M(m-1,m)\).

\begin{example}[First minimal-model cases]
\label{ex:first-phi13-minimal-flows}
For \(m=3\), \(\mathcal M(3,4)\) is the Ising CFT.  The label
\((1,3)\) is identified with \((2,1)\), and
\[
  h_{1,3}^{(3)}=\frac12,\qquad \Delta_{1,3}^{(3)}=1.
\]
This is the thermal perturbation, whose massive scaling limit is the free
massive Majorana theory discussed in
Chapter~\ref{ch:integrable-thermodynamic-bethe-ansatz}.  For \(m=4\),
\[
  c_4=\frac{7}{10},\qquad c_3=\frac12,\qquad c_4-c_3=\frac15,
\]
which is the central-charge drop of the tricritical-Ising to Ising massless
branch.  These statements fix the conformal arithmetic; the sign of the
coupling and the operator normalization still have to be declared before
comparing with scattering or TBA conventions.
\end{example}

\section{Finite-Size Scaling Function}

Let \(E_0(R)\) be the ground-state energy on a spatial circle of
circumference \(R\).  Define
\[
  c_{\mathrm{eff}}(R)
  =
  -\frac{6R}{\pi}E_0(R).
\]
For a CFT with periodic boundary conditions and cylinder Hamiltonian
\[
  H_R=\frac{2\pi}{R}\left(L_0+\bar L_0-\frac{c}{12}\right),
\]
the vacuum energy is
\[
  E_0(R)=-\frac{\pi c}{6R},
\]
so \(c_{\mathrm{eff}}(R)=c\).  In a nonunitary CFT the cylinder ground state
may be a primary of negative weight; then the same formula gives
\[
  c_{\mathrm{eff}}=c-24h_{\min},
\]
where \(h_{\min}\) is the lowest diagonal conformal weight.  This is the
convention used for the Lee--Yang example in
Chapter~\ref{ch:integrable-thermodynamic-bethe-ansatz}.

For a massive infrared theory with a unique vacuum,
\[
  E_0(R)=\mathcal E_{\mathrm{bulk}}R+o(R^0)
  \qquad (R\to\infty).
\]
The bulk energy density \(\mathcal E_{\mathrm{bulk}}\) is a local extensive
coordinate.  Hence comparisons of ultraviolet and infrared central charges
must say whether the extensive term has been subtracted.  The subtracted
quantity is
\[
  E_0^{\mathrm{sub}}(R)=E_0(R)-\mathcal E_{\mathrm{bulk}}R,
  \qquad
  c_{\mathrm{sub}}(R)=-\frac{6R}{\pi}E_0^{\mathrm{sub}}(R).
\]

\begin{definition}[Finite-size identification datum]
\label{def:finite-size-identification-datum}
A finite-size identification between a perturbed CFT and a factorized
scattering model consists of:
\begin{enumerate}
\item a Hilbert space on the circle and a self-adjoint Hamiltonian \(H_R\) for
each \(R\) in the scaling regime;
\item a normalization of the vacuum energy, including the value of
\(\mathcal E_{\mathrm{bulk}}\);
\item a mass convention for the infinite-volume particles;
\item a proof, or an explicitly stated hypothesis, that the large-volume
mirror-channel thermodynamics computes \(E_0(R)\).
\end{enumerate}
With these data declared, the equality between a TBA scaling function and a
QFT finite-size energy is a precise mathematical assertion.
\end{definition}

\section{Trace Sum Rule as a Local Test of the Flow}

Finite-size central charges are necessary but not sufficient data for
identifying a local RG trajectory.  A sharper test uses the trace of the
stress tensor.  In this section \(\Theta=T^\mu{}_\mu\) denotes the flat-space
Cartesian trace in the stress-tensor normalization for which the
Zamolodchikov sum rule below has coefficient \(3/(4\pi)\).  If a different
normalization of the stress tensor or the perturbing source is used, the same
statements hold after replacing \(\Theta\) by the correspondingly normalized
trace operator.

\begin{definition}[Zamolodchikov sum-rule datum]
\label{def:zamolodchikov-sum-rule-datum}
A two-dimensional Euclidean RG trajectory has a Zamolodchikov sum-rule datum
if the following data are specified.
\begin{enumerate}
\item A reflection-positive local QFT on \(\mathbb R^2\) with a symmetric
conserved stress tensor \(T_{\mu\nu}\), trace
\(\Theta=T^\mu{}_\mu\), and connected trace two-point distribution
\[
  G_\Theta(x)=
  \langle \Theta(x)\Theta(0)\rangle_{\rm c}.
\]
\item A distributional extension of \(G_\Theta\) to \(x=0\) whose contact
terms are fixed by the stress-tensor source chart.  The weighted pairing
\[
  \int_{\mathbb R^2} d^2x\, |x|^2 G_\Theta(x)
\]
is required to be finite and independent of the remaining allowed
point-supported contact terms.  This condition is automatic for contact terms
annihilated by the test function \(|x|^2\), but it must be checked when
derivatives of delta functions are permitted by the source scheme.
\item A continuous function \(C(R)\), \(R>0\), whose endpoint limits exist,
\[
  \lim_{R\downarrow0}C(R)=c_{\rm UV},
  \qquad
  \lim_{R\to\infty}C(R)=c_{\rm IR},
\]
and whose radial derivative is represented by
\begin{equation}
  \frac{dC(R)}{d\log R}
  =
  -\frac{3}{4\pi}
  \int_0^{2\pi} d\alpha\,
  R^4\,G_\Theta(R e^{\ii\alpha})
  \label{eq:zamolodchikov-radial-derivative-datum}
\end{equation}
in the distributional sense after smearing in \(R\).
\end{enumerate}
\end{definition}

The last item is the local Ward-identity content of the two-dimensional
\(c\)-theorem in the present normalization.  It is not a definition of
\(G_\Theta\); it is a theorem-level property to be established from stress
tensor conservation, Euclidean covariance, positivity, and the regularity
assumptions of the chosen QFT framework.

\begin{proposition}[Trace sum rule]
\label{prop:integrable-flow-trace-sum-rule}
For a trajectory with a Zamolodchikov sum-rule datum,
\begin{equation}
  c_{\rm UV}-c_{\rm IR}
  =
  \frac{3}{4\pi}
  \int_{\mathbb R^2} d^2x\, |x|^2
  \langle \Theta(x)\Theta(0)\rangle_{\rm c}.
  \label{eq:integrable-flow-trace-sum-rule}
\end{equation}
If reflection positivity makes \(G_\Theta(x)\) a nonnegative radial spectral
density at separated points, then \(c_{\rm UV}\geq c_{\rm IR}\).
\end{proposition}

\begin{proof}
Integrating \eqref{eq:zamolodchikov-radial-derivative-datum} from
\(R=\epsilon\) to \(R=L\) gives
\[
  C(\epsilon)-C(L)
  =
  \frac{3}{4\pi}
  \int_\epsilon^L \frac{dR}{R}
  \int_0^{2\pi}d\alpha\,
  R^4G_\Theta(R e^{\ii\alpha}).
\]
Since \(d^2x=R\,dR\,d\alpha\), the right side is
\[
  \frac{3}{4\pi}
  \int_{\epsilon<|x|<L}d^2x\, |x|^2G_\Theta(x).
\]
The integrability hypothesis in
Definition~\ref{def:zamolodchikov-sum-rule-datum} permits the limits
\(\epsilon\downarrow0\) and \(L\to\infty\), while the endpoint hypothesis
identifies the left side with \(c_{\rm UV}-c_{\rm IR}\).  Positivity of the
weighted trace two-point distribution gives the inequality.
\end{proof}

\begin{proposition}[Form-factor version of the trace sum rule]
\label{prop:form-factor-trace-sum-rule}
Consider a massive integrable trajectory with particle species \(a\), masses
\(m_a>0\), and trace-operator form factors
\[
  F^{\Theta}_{a_1\cdots a_n}(\theta_1,\ldots,\theta_n)
  =
  \langle 0|\Theta(0)|
  \theta_1,a_1;\ldots;\theta_n,a_n\rangle_{\rm in}.
\]
Assume that the connected Euclidean two-point function has the absolutely
convergent spectral expansion
\begin{equation}
  G_\Theta(x)
  =
  \sum_{n\geq1}\frac1{n!}
  \sum_{a_1,\ldots,a_n}
  \int_{\mathbb R^n}\prod_{j=1}^n\frac{d\theta_j}{2\pi}\,
  \left|F^{\Theta}_{a_1\cdots a_n}(\theta_1,\ldots,\theta_n)\right|^2
  e^{-|x|\sum_j m_{a_j}\cosh\theta_j}
  \label{eq:trace-ff-spectral-expansion}
\end{equation}
and that Fubini's theorem applies to the \(|x|^2\)-weighted integral.  Then
\begin{equation}
  c_{\rm UV}-c_{\rm IR}
  =
  9
  \sum_{n\geq1}\frac1{n!}
  \sum_{a_1,\ldots,a_n}
  \int_{\mathbb R^n}\prod_{j=1}^n\frac{d\theta_j}{2\pi}\,
  \frac{
  \left|F^{\Theta}_{a_1\cdots a_n}(\theta_1,\ldots,\theta_n)\right|^2}
  {\left(\sum_j m_{a_j}\cosh\theta_j\right)^4}.
  \label{eq:trace-form-factor-c-sum-rule}
\end{equation}
\end{proposition}

\begin{proof}
Insert \eqref{eq:trace-ff-spectral-expansion} into
\eqref{eq:integrable-flow-trace-sum-rule}.  Absolute convergence permits the
exchange of the spacetime integral with the rapidity integrals and species
sums.  For fixed rapidities set
\[
  E(\theta_1,\ldots,\theta_n)=\sum_j m_{a_j}\cosh\theta_j>0 .
\]
The radial integral is elementary:
\[
  \frac{3}{4\pi}
  \int_{\mathbb R^2}d^2x\,|x|^2e^{-E|x|}
  =
  \frac{3}{4\pi}(2\pi)
  \int_0^\infty d\rho\,\rho^3e^{-E\rho}
  =
  \frac{3}{2}\frac{3!}{E^4}
  =
  \frac9{E^4}.
\]
Substitution gives the displayed formula.
\end{proof}

\begin{example}[\(\phi_{1,3}\) minimal-model flow target]
\label{ex:phi13-trace-sum-rule-target}
For the massless branch expected to connect
\(\mathcal M(m,m+1)\) to \(\mathcal M(m-1,m)\), any proposed trace form
factors and normalization must satisfy the numerical target
\[
  9
  \sum_{n\geq1}\frac1{n!}
  \sum_{a_1,\ldots,a_n}
  \int\prod_{j=1}^n\frac{d\theta_j}{2\pi}\,
  \frac{|F^\Theta_{a_1\cdots a_n}(\theta_1,\ldots,\theta_n)|^2}
  {(\sum_j m_{a_j}\cosh\theta_j)^4}
  =
  \frac{12}{m(m^2-1)} .
\]
The right side is the central-charge drop proved in
Proposition~\ref{prop:phi13-minimal-model-data}.  This equality is stronger
than matching the endpoint values of a TBA scaling function: it tests the
local normalization of \(\Theta\), the form-factor solution, and the
short-distance-to-long-distance interpolation of the trajectory.
\end{example}

\section{TBA Scaling Function}

For a diagonal factorized massive theory, use the kernel convention of
Chapter~\ref{ch:integrable-thermodynamic-bethe-ansatz},
\[
  \varphi_{ab}(\theta)
  =
  \frac{1}{2\pi\ii}\frac{d}{d\theta}\log S_{ab}(\theta).
\]
Let \(m_a=\mu_a M\), where \(M\) is the reference mass and
\(\mu_a>0\).  The pseudoenergies are solutions of
\[
  \epsilon_a(\theta;r)
  =
  r\,\mu_a\cosh\theta
  -\sum_b\int_{\mathbb R}d\theta'\,
  \varphi_{ab}(\theta-\theta')
  \log\!\left(1+e^{-\epsilon_b(\theta';r)}\right),
  \qquad r=MR .
\]
With the bulk term separated, the mirror-channel formula gives
\[
  E_0^{\mathrm{sub}}(R)
  =
  -\frac{M}{2\pi}
  \sum_a\mu_a\int_{\mathbb R}d\theta\,
  \cosh\theta\,
  \log\!\left(1+e^{-\epsilon_a(\theta;r)}\right).
\]
Consequently
\[
  c_{\mathrm{TBA}}(r)
  =
  \frac{3r}{\pi^2}
  \sum_a\mu_a\int_{\mathbb R}d\theta\,
  \cosh\theta\,
  \log\!\left(1+e^{-\epsilon_a(\theta;r)}\right).
\]
This formula is a theorem about the variational TBA functional under the
hypotheses of Chapter~\ref{ch:integrable-thermodynamic-bethe-ansatz}; its
identification with the finite-size vacuum energy of a local QFT requires the
additional datum in Definition~\ref{def:finite-size-identification-datum}.

\begin{controlledapproximation}[Plateau reduction]
\label{ca:tba-plateau-reduction}
Assume that as \(r\downarrow0\) the solution has a plateau interval whose
width tends to infinity, that the boundary layers at the ends of the plateau
contribute only the standard Rogers-dilogarithm endpoints, and that the
convolution kernels may be replaced on the plateau by their total integrals
\[
  N_{ab}=\int_{\mathbb R}d\theta\,\varphi_{ab}(\theta).
\]
The plateau values \(Y_a=\exp(\epsilon_a)\) then obey
\[
  \log Y_a=-\sum_b N_{ab}\log(1+Y_b^{-1}).
\]
The corresponding ultraviolet contribution to the effective central charge is
\[
  c_{\mathrm{UV}}
  =
  \frac{6}{\pi^2}\sum_a
  L\!\left(\frac{1}{1+Y_a}\right),
\]
where
\[
  L(x)
  =
  -\frac12\int_0^x dt\,
  \left(
  \frac{\log(1-t)}{t}
  +\frac{\log t}{1-t}
  \right)
  =
  \operatorname{Li}_2(x)+\frac12\log x\log(1-x)
\]
for \(0<x<1\).
\end{controlledapproximation}

\begin{proof}
The displayed algebraic equation is obtained by dropping the driving term on
the plateau and replacing each convolution by the total kernel integral.  To
derive the dilogarithm contribution, differentiate the TBA equation with
respect to rapidity in a boundary layer where the pseudoenergy changes from a
constant plateau value to its asymptotic value.  The same integration by
parts used in the Lee--Yang calculation of
Chapter~\ref{ch:integrable-thermodynamic-bethe-ansatz} reduces the scaling
integral for species \(a\) to
\[
  \frac{6}{\pi^2}
  \int d\epsilon_a\,
  \log(1+e^{-\epsilon_a})
  \frac{d}{d\epsilon_a}
  \log(1+e^{\epsilon_a})^{-1}.
\]
With \(x=(1+e^{\epsilon_a})^{-1}\), this is the Rogers dilogarithm evaluated
between the endpoint values.  Under the present plateau hypothesis the
nontrivial endpoint is \(x_a=(1+Y_a)^{-1}\), giving the stated sum.  The
argument is a controlled asymptotic reduction.  A complete application to a
particular model includes a proof of the stated plateau hypothesis for that
model.
\end{proof}

\section{Massless Flows}

An integrable massless flow contains right- and left-moving asymptotic
particles in the infrared scaling description.  A right mover is
parametrized by
\[
  E=p=\frac{M}{2}e^\theta ,
\]
and a left mover by
\[
  E=-p=\frac{M}{2}e^{-\theta}.
\]
The parameter \(M\) is the crossover scale.  The TBA now has right-right,
left-left, and right-left kernels.  The right-left kernel records the
interaction between the two chiral sectors along the flow; its normalization
controls the interpolation between the ultraviolet and infrared plateaux.

\begin{definition}[Massless factorized-flow datum]
\label{def:massless-factorized-flow-datum}
A diagonal massless factorized-flow datum consists of:
\begin{enumerate}
\item right species \(a\in A_R\) and left species \(\alpha\in A_L\);
\item analytic two-body amplitudes
\[
  S^{RR}_{ab}(\theta),\qquad
  S^{LL}_{\alpha\beta}(\theta),\qquad
  S^{RL}_{a\alpha}(\theta)
\]
obeying the declared unitarity, crossing, and bootstrap equations in their
analytic strips;
\item a TBA system with driving terms
\[
  r\,\mu_a e^\theta/2,\qquad r\,\nu_\alpha e^{-\theta}/2,
\]
and kernels obtained from the logarithmic derivatives of the amplitudes in
the same sign convention;
\item ultraviolet and infrared local-operator charts to which the finite-size
scaling function and form factors are to be matched.
\end{enumerate}
\end{definition}

For the minimal-model \(\phi_{1,3}\) massless branch, the expected endpoint
central charges satisfy
\[
  \lim_{r\downarrow0}c_{\mathrm{sub}}(r)=c_m,\qquad
  \lim_{r\to\infty}c_{\mathrm{sub}}(r)=c_{m-1}.
\]
The value of \(c_m-c_{m-1}\) is fixed by
Proposition~\ref{prop:phi13-minimal-model-data}.  The exact TBA system must
reproduce these two limits after the bulk term and mass scale have been
matched.  This central-charge comparison is one component of the
identification; the full identification also includes the local operator
algebras and their OPE data in the ultraviolet and infrared.

\section{Operator Data Along the Flow}

The local-correlator part of the flow is supplied by form factors and their
short-distance normalization.  For a local operator
\(\widehat{\mathcal A}\), define its infinite-volume form factors by
\[
  F^{\mathcal A}_{a_1\cdots a_n}(\theta_1,\ldots,\theta_n)
  =
  \langle 0|\widehat{\mathcal A}(0)|
  \theta_1,a_1;\ldots;\theta_n,a_n\rangle_{\mathrm{in}},
\]
where the ordering convention and particle-state normalization are those of
Chapter~\ref{ch:integrable-form-factor-bootstrap-local-operators}.  The
ultraviolet scaling dimension is tested by the short-distance behavior
\[
  \langle \widehat{\mathcal A}(x)\widehat{\mathcal A}(0)\rangle_{\mathrm c}
  \sim
  \frac{C_{\mathcal A}}{|x|^{2\Delta_{\mathcal A}}}
  \qquad (M|x|\to0),
\]
with \(C_{\mathcal A}\) fixed by a source convention.  The perturbing operator
itself must also satisfy the trace relation in the chosen normalization:
\[
  \Theta(x)
  =
  \kappa_\Theta\,y_{\mathcal O}\,g\,\mathcal O(x)
  +O(g^2)
\]
when the Euclidean action is perturbed as
\(S_{\mathcal C}+g\int d^2x\,\mathcal O(x)\) and the stress tensor is
normalized by \(T^\mu{}_\mu=\Theta\).  The constant \(\kappa_\Theta\) is fixed
by the Weyl-variation convention for the stress tensor and by whether a factor
such as \(1/(2\pi)\) is included in the source term.  A chapter using this
relation states \(\kappa_\Theta\) before using numerical trace-normalization
formulas.

\begin{proposition}[Necessary matching conditions for an integrable flow]
\label{prop:necessary-integrable-flow-matching}
Suppose a factorized scattering model is identified with a relevant
perturbation of \(\mathcal C\) by \(\mathcal O\) in the sense of
Definition~\ref{def:renormalized-perturbed-cft-chart}.  Then the following
data must be mutually compatible:
\begin{enumerate}
\item the particle masses and scattering amplitudes determine a TBA scaling
function whose ultraviolet limit equals the effective central charge of
\(\mathcal C\);
\item if the infrared limit is conformal, the large-\(R\) subtracted
finite-size energy gives the infrared central charge;
\item the form-factor spectral expansion of the operator corresponding to
\(\mathcal O\) has the ultraviolet two-point normalization fixed in the
perturbed-CFT chart;
\item the trace operator extracted from the finite-size response to a change
of \(R\) agrees, to first order in \(g\), with the perturbing operator through
the displayed trace relation.
\end{enumerate}
\end{proposition}

\begin{proof}
The first two assertions are consequences of the definition of
\(c_{\mathrm{sub}}(R)\) and of the CFT cylinder Hamiltonian.  A mismatch of
central charge would contradict equality of the finite-size vacuum energies
in either scaling limit.  The third assertion follows because the source
coordinate in Definition~\ref{def:renormalized-perturbed-cft-chart} fixes the
normalization of \(\mathcal O\); a different form-factor normalization would
compute a different source derivative of the Schwinger functional.  For the
fourth assertion, vary the perturbed action under a Weyl rescaling.  The
integrated insertion has scaling defect \(y_{\mathcal O}\), and the declared
stress-tensor normalization converts the Weyl variation of the action into
the trace insertion.  Higher powers of \(g\) are contact-term and operator
mixing data of the same renormalized chart.
\end{proof}

\begin{openproblem}[Rigorous integrable-flow reconstruction]
\label{op:integrable-flow-reconstruction}
Construct a theorem that starts from factorized scattering data, form-factor
solutions satisfying locality and convergence estimates, and TBA finite-size
data, and reconstructs a local two-dimensional QFT whose ultraviolet limit is
a specified CFT perturbation.  The theorem should identify the operator
topology in which the ultraviolet limit is taken, prove the finite-size
Hamiltonian/TBA equality, and show that the OPE coefficients obtained from
short-distance limits agree with the declared CFT data.
\end{openproblem}
