\chapter{Nested Bethe Ansatz and Matrix Bethe--Yang Equations}
\label{ch:integrable-nested-bethe-ansatz-matrix-bethe-yang}

\ConventionsLorentzian

The algebraic Bethe ansatz of
Chapter~\ref{ch:integrable-algebraic-bethe-ansatz-transfer-matrices}
diagonalizes a transfer matrix by turning the Yang--Baxter equation into
commuting finite-dimensional operators.  This chapter develops the
higher-rank version needed for non-diagonal finite-volume scattering.  The
physical rapidities are fixed by Bethe--Yang quantization, while auxiliary
Bethe roots diagonalize the internal transfer matrix.  The auxiliary roots
are not extra particles: they are spectral coordinates for the internal
wavefunction of a multiparticle state.

\section{Matrix Bethe--Yang Quantization}

Let a massive integrable QFT have \(n\) stable particles with rapidities
\(\theta_1,\ldots,\theta_n\), masses \(m_{a_j}\), and internal species spaces
\(V_{a_j}\).  Write
\[
  \mathcal V_{\boldsymbol a}
  =
  V_{a_1}\otimes\cdots\otimes V_{a_n}.
\]
For a chosen particle \(j\), transporting it once around a spatial circle of
circumference \(L\) produces the phase
\(\exp(\ii m_{a_j}L\sinh\theta_j)\) and the ordered product of two-body
scatterings through the other particles.  In the \(R\)-matrix convention,
this ordered product is an internal transfer matrix
\[
  \mathcal T_j(\theta_j\mid\theta_1,\ldots,\widehat{\theta_j},\ldots,\theta_n)
  \in \operatorname{End}(\mathcal V_{\boldsymbol a}).
\]
The hat means that \(\theta_j\) is omitted from the list of spectator
rapidities.  The Yang--Baxter equation implies that the different
\(\mathcal T_j\)'s commute on the common meromorphic domain after the
ordering convention for identical particles has been fixed.

\begin{definition}[Matrix Bethe--Yang state]
\label{def:matrix-bethe-yang-state}
A finite-volume \(n\)-particle matrix Bethe--Yang state is a vector
\(\Psi\in\mathcal V_{\boldsymbol a}\) and rapidities
\(\theta_1,\ldots,\theta_n\) such that \(\Psi\) is a simultaneous eigenvector
of all internal transfer matrices,
\[
  \mathcal T_j\Psi=\tau_j\Psi ,
\]
and
\begin{equation}
  \exp\!\left(\ii m_{a_j}L\sinh\theta_j\right)\tau_j=1,
  \qquad j=1,\ldots,n .
  \label{eq:matrix-bethe-yang}
\end{equation}
\end{definition}

When the scattering is diagonal, \(\tau_j\) is the product of scalar
two-particle phases and \eqref{eq:matrix-bethe-yang} reduces to the
Bethe--Yang equation in
Chapter~\ref{ch:integrable-thermodynamic-bethe-ansatz}.  In the non-diagonal
case the eigenvalue \(\tau_j\) is obtained by an auxiliary Bethe ansatz.

\section{RTT Algebra and Yangian Evaluation Data}

Let \(V=\C^N\) and let \(E_{ab}\in\operatorname{End}(V)\) be the elementary
matrix with \(E_{ab}e_c=\delta_{bc}e_a\).  For the rational \(SU(N)\) chain
we use
\[
  R(u)=u\,\id+\ii P,\qquad
  P=\sum_{a,b=1}^{N}E_{ab}\otimes E_{ba}.
\]
The monodromy
\[
  T_0(u)=R_{0L}(u-\xi_L)\cdots R_{01}(u-\xi_1)
\]
is written as
\[
  T_0(u)=\sum_{a,b=1}^{N}E_{ab}^{(0)}\otimes T_{ab}(u),
  \qquad T_{ab}(u)\in\operatorname{End}(\mathcal H_L),
\]
where \(\mathcal H_L=(\C^N)^{\otimes L}\).  The RTT relation
\[
  R_{00'}(u-v)T_0(u)T_{0'}(v)
  =
  T_{0'}(v)T_0(u)R_{00'}(u-v)
\]
is the defining finite-dimensional representation of the rational Yangian
RTT algebra.  Equivalently, the matrix entries obey the component relation
\begin{equation}
  (u-v)[T_{ab}(u),T_{cd}(v)]
  =
  \ii\bigl(
    T_{cb}(v)T_{ad}(u)-T_{cb}(u)T_{ad}(v)
  \bigr).
  \label{eq:rational-yangian-rtt-components}
\end{equation}

For \(R(u)=u+\ii P\), the operator identity RTT is equivalent to
\eqref{eq:rational-yangian-rtt-components} for all
\(a,b,c,d=1,\ldots,N\).
Insert \(P=\sum_{p,q}E_{pq}^{(0)}E_{qp}^{(0')}\) and
\(T_0(u)=\sum E_{ab}^{(0)}T_{ab}(u)\),
\(T_{0'}(v)=\sum E_{cd}^{(0')}T_{cd}(v)\) into RTT.  The coefficient of
\(E_{ab}^{(0)}E_{cd}^{(0')}\) in the term proportional to \(u-v\) is
\((u-v)(T_{ab}(u)T_{cd}(v)-T_{cd}(v)T_{ab}(u))\).  The coefficient in the
permutation terms is
\(\ii(T_{cb}(v)T_{ad}(u)-T_{cb}(u)T_{ad}(v))\).  Equality of the coefficients
is exactly \eqref{eq:rational-yangian-rtt-components}, and the converse is
obtained by summing these coefficient identities.

Expanding \(T_{ab}(u)\) for large \(u\) gives level-zero and higher Yangian
charges.  For a finite chain the representation is not an additional
postulate: it is the evaluation representation produced by the concrete
monodromy above.  The Yangian language is nevertheless useful because the
nested roots are organized by the simple roots of \(\mathfrak{sl}_N\), and
the recursive auxiliary monodromy below is the same RTT construction applied
to smaller residual color spaces.

\section{Recursive Nested Monodromy}

Choose the reference vector
\[
  \Omega_L=e_1^{\otimes L}.
\]
With the shifted \(L\)-operator convention of
Chapter~\ref{ch:integrable-algebraic-bethe-ansatz-transfer-matrices},
\[
  T_{11}(u)\Omega_L=\left(u+\frac{\ii}{2}\right)^L\Omega_L,
  \qquad
  T_{aa}(u)\Omega_L=\left(u-\frac{\ii}{2}\right)^L\Omega_L
  \quad(a=2,\ldots,N),
\]
and the lower-left entries annihilate \(\Omega_L\).  The first-level
creation operators are the entries \(B_\alpha(u)=T_{1,\alpha+1}(u)\),
\(\alpha=1,\ldots,N-1\).  A first-level Bethe vector has the form
\[
  \Psi^{(1)}
  =
  \sum_{\alpha_1,\ldots,\alpha_{M_1}=1}^{N-1}
  F^{\alpha_1\cdots\alpha_{M_1}}\,
  B_{\alpha_1}(u_1^{(1)})\cdots B_{\alpha_{M_1}}(u_{M_1}^{(1)})\Omega_L .
\]
The coefficient tensor \(F\) is not arbitrary.  RTT relations among the
\(B_\alpha\)'s show that exchanging two first-level roots acts on their color
indices by the rational \(R\)-matrix on \(\C^{N-1}\).  Therefore the
unwanted terms in the action of the full transfer matrix are controlled by
an auxiliary \(SU(N-1)\) transfer matrix on
\((\C^{N-1})^{\otimes M_1}\).  This is the origin of nesting.

\begin{construction}[Nested auxiliary monodromy]
\label{con:nested-auxiliary-monodromy}
Let \(V^{(r)}\simeq\C^{N-r}\) be the residual color space at nesting level
\(r\), with \(r=1,\ldots,N-1\).  Put \(M_0=L\), and let
\(\{u_j^{(0)}\}\) denote the physical inhomogeneities in the spin-chain
problem or the physical rapidities after conversion to the rational
auxiliary variable.  Given roots
\(\{u_j^{(r-1)}\}_{j=1}^{M_{r-1}}\), define the level-\(r\) auxiliary
monodromy by
\[
  T^{(r)}_0(u)
  =
  R^{(N-r+1)}_{0M_{r-1}}(u-u_{M_{r-1}}^{(r-1)})
  \cdots
  R^{(N-r+1)}_{01}(u-u_1^{(r-1)}),
\]
where \(R^{(m)}(u)=u+\ii P\) acts on \(\C^m\otimes\C^m\).  The level-\(r\)
transfer matrix is \(t^{(r)}(u)=\operatorname{tr}_0T^{(r)}_0(u)\).  The roots
\(\{u_j^{(r)}\}_{j=1}^{M_r}\) are the Bethe roots used to diagonalize
\(t^{(r)}\) by the same pseudovacuum construction, now with residual rank
smaller by one.
\end{construction}

\begin{theorem}[Recursive origin of the nested equations]
\label{thm:recursive-nested-bethe-origin}
For the rational fundamental \(SU(N)\) chain with generic inhomogeneities and
generic roots, the repeated construction
\ref{con:nested-auxiliary-monodromy} gives an eigenvector of the transfer
matrix whenever the roots satisfy the pole-cancellation equations
\begin{equation}
  \left(
    \frac{u_j^{(r)}+\frac{\ii}{2}\delta_{r1}}
         {u_j^{(r)}-\frac{\ii}{2}\delta_{r1}}
  \right)^L
  =
  \prod_{s=1}^{N-1}
  \prod_{\substack{k=1\\(s,k)\ne(r,j)}}^{M_s}
  \frac{u_j^{(r)}-u_k^{(s)}+\frac{\ii}{2}A_{rs}}
       {u_j^{(r)}-u_k^{(s)}-\frac{\ii}{2}A_{rs}},
  \label{eq:su-n-nested-bethe-cartan}
\end{equation}
where \(A_{rs}=2\delta_{rs}-\delta_{r,s+1}-\delta_{r,s-1}\) is the
\(A_{N-1}\) Cartan matrix.  The energy in the nearest-neighbor rational
Hamiltonian normalization depends only on first-level roots:
\begin{equation}
  E=\sum_{j=1}^{M_1}\frac{1}{(u_j^{(1)})^2+1/4}.
  \label{eq:su-n-rational-energy-first-level}
\end{equation}
\end{theorem}

\begin{proof}
The proof is an induction on \(N\).  The case \(N=2\) is precisely
Theorem~\ref{thm:aba-xxx-bethe}: the transfer eigenvalue is meromorphic with
possible simple poles at first-level roots, and the residue at
\(u=u_j^{(1)}\) vanishes exactly when the \(SU(2)\) Bethe equation holds.

Assume the statement for residual rank \(N-r\).  At level \(r\), the RTT
commutation relations move the diagonal entries of \(T^{(r)}(u)\) through the
creation operators \(B^{(r)}_\alpha(u_j^{(r)})\).  The wanted term is the one
in which no spectral parameter is exchanged.  The unwanted terms replace one
creation operator by \(B^{(r)}_\alpha(u)\) and carry coefficients that are
matrix elements of the next auxiliary transfer matrix \(t^{(r+1)}\) on
\((V^{(r+1)})^{\otimes M_r}\).  By the induction hypothesis those matrix
coefficients are diagonal exactly when the roots at levels
\(r+1,\ldots,N-1\) obey their own pole-cancellation equations.

Only three kinds of factors can appear in the residue at a root
\(u_j^{(r)}\).  First, exchanging \(u_j^{(r)}\) with another root at the same
level uses the two-body rational \(R\)-matrix on the same residual color
space and contributes the shifts \(\pm\ii\), which are the entries
\(\pm\ii A_{rr}/2\).  Second, the next auxiliary transfer matrix contributes
only adjacent Dynkin-node factors, because after one nesting step the root
\(u^{(r)}\) is an inhomogeneity for the \(SU(N-r)\) residual problem; this
gives the shifts \(\mp\ii/2\) for \(s=r\pm1\), namely
\(\pm\ii A_{rs}/2\).  Third, the vacuum diagonal eigenvalues of the original
monodromy enter only at \(r=1\), giving the source factor
\(((u+\ii/2)/(u-\ii/2))^L\).  Non-adjacent nodes correspond to commuting
residual color sectors and give \(A_{rs}=0\), hence the factor \(1\).
Setting every residue to zero is exactly
\eqref{eq:su-n-nested-bethe-cartan}.  The Hamiltonian is the logarithmic
derivative of the first-level transfer matrix at the regular point, so the
same calculation as in the \(SU(2)\) case gives
\eqref{eq:su-n-rational-energy-first-level}; auxiliary levels select the
internal eigenvector but carry no direct nearest-neighbor energy.
\end{proof}

Expanding \eqref{eq:su-n-nested-bethe-cartan} gives the familiar displayed
form
\begin{equation}
\begin{aligned}
  \left(
    \frac{u_j^{(1)}+\ii/2}{u_j^{(1)}-\ii/2}
  \right)^L
  &=
  \prod_{\substack{k=1\\ k\neq j}}^{M_1}
  \frac{u_j^{(1)}-u_k^{(1)}+\ii}
       {u_j^{(1)}-u_k^{(1)}-\ii}
  \prod_{k=1}^{M_2}
  \frac{u_j^{(1)}-u_k^{(2)}-\ii/2}
       {u_j^{(1)}-u_k^{(2)}+\ii/2},
\\
  1
  &=
  \prod_{\substack{k=1\\ k\neq j}}^{M_r}
  \frac{u_j^{(r)}-u_k^{(r)}+\ii}
       {u_j^{(r)}-u_k^{(r)}-\ii}
  \prod_{\sigma=\pm1}
  \prod_{k=1}^{M_{r+\sigma}}
  \frac{u_j^{(r)}-u_k^{(r+\sigma)}-\ii/2}
       {u_j^{(r)}-u_k^{(r+\sigma)}+\ii/2},
\\[-0.4em]
  &\hspace{19em} r=2,\ldots,N-2,
\\
  1
  &=
  \prod_{\substack{k=1\\ k\neq j}}^{M_{N-1}}
  \frac{u_j^{(N-1)}-u_k^{(N-1)}+\ii}
       {u_j^{(N-1)}-u_k^{(N-1)}-\ii}
  \prod_{k=1}^{M_{N-2}}
  \frac{u_j^{(N-1)}-u_k^{(N-2)}-\ii/2}
       {u_j^{(N-1)}-u_k^{(N-2)}+\ii/2}.
\end{aligned}
  \label{eq:su-n-nested-bethe}
\end{equation}
Products over an empty set are \(1\).

\section{Dressed-Vacuum Transfer Eigenvalue}

The nested equations are most useful when one keeps track of the transfer
eigenvalue that produced them.  We record the expression in a gauge in which
the pole-cancellation calculation is transparent.  Let
\[
  Q_r(u)=\prod_{j=1}^{M_r}(u-u_j^{(r)}),\qquad
  Q_0(u)=Q_N(u)=1,
\]
and use the half-shift notation
\[
  f^{[k]}(u)=f\!\left(u+\frac{\ii k}{2}\right).
\]
For the homogeneous fundamental chain define
\[
  \phi_1(u)=(u+\ii)^L,\qquad
  \phi_a(u)=u^L\quad(a=2,\ldots,N).
\]
With a diagonal twist \(g=\operatorname{diag}(g_1,\ldots,g_N)\) the
fundamental transfer eigenvalue can be written, after an overall scalar
normalization and a harmless shift of the spectral coordinate, as
\begin{equation}
  \Lambda(u)
  =
  \sum_{a=1}^{N}
  g_a\,\phi_a(u)\,
  \frac{Q_{a-1}^{[a+1]}(u)}{Q_{a-1}^{[a-1]}(u)}
  \frac{Q_a^{[a-2]}(u)}{Q_a^{[a]}(u)} .
  \label{eq:su-n-dressed-vacuum-eigenvalue}
\end{equation}
This is the dressed-vacuum form of the eigenvalue.  The adjective means only
that the vacuum diagonal eigenvalues \(\phi_a\) are multiplied by rational
``dressing'' factors built from the \(Q\)-functions; it is not an additional
physical vacuum.

\paragraph{Pole-cancellation derivation of the twisted nested equations.}
Assume the roots are generic and the twist eigenvalue ratios are finite.  The
meromorphic function \eqref{eq:su-n-dressed-vacuum-eigenvalue} has no pole at
the would-be pole associated with \(u_j^{(r)}\) precisely when
\begin{equation}
  \frac{g_r}{g_{r+1}}
  \frac{\phi_r(u_j^{(r)}-\frac{\ii r}{2})}
       {\phi_{r+1}(u_j^{(r)}-\frac{\ii r}{2})}
  =
  \prod_{s=1}^{N-1}
  \prod_{\substack{k=1\\(s,k)\ne(r,j)}}^{M_s}
  \frac{u_j^{(r)}-u_k^{(s)}+\frac{\ii}{2}A_{rs}}
       {u_j^{(r)}-u_k^{(s)}-\frac{\ii}{2}A_{rs}} .
  \label{eq:twisted-su-n-dva-bethe}
\end{equation}
For the untwisted homogeneous chain this reduces to
\eqref{eq:su-n-nested-bethe-cartan}.
This is a residue calculation inside the dressed-vacuum formula, not an
independent theorem: it is the local algebraic step that converts
polynomiality of the transfer eigenvalue into the nested Bethe equations.

The root \(u_j^{(r)}\) appears as a pole of \(\Lambda(u)\) at the common
point \(u=u_j^{(r)}-\ii r/2\) in precisely two neighboring summands,
namely the \(a=r\) and \(a=r+1\) summands.  All other summands are regular
there because their denominator \(Q\)-functions have different node labels.
Multiplying the residue equation by the nonzero common factors gives
\[
\begin{aligned}
0
&=
g_r\phi_r
  \frac{Q_{r-1}^{[r+1]}}{Q_{r-1}^{[r-1]}}
  Q_r^{[r-2]}
\\
&\quad+
g_{r+1}\phi_{r+1}
  Q_r^{[r+2]}
  \frac{Q_{r+1}^{[r-1]}}{Q_{r+1}^{[r+1]}}
\end{aligned}
\]
evaluated at \(u=u_j^{(r)}-\ii r/2\); the omitted common denominator is the
same derivative \(Q_r'(u_j^{(r)})\).  Solving for the ratio of the two vacuum
terms gives
\[
  \frac{g_r\phi_r}{g_{r+1}\phi_{r+1}}
  =
  -\frac{Q_r(u_j^{(r)}+\ii)}{Q_r(u_j^{(r)}-\ii)}
  \frac{Q_{r-1}(u_j^{(r)}-\ii/2)}
       {Q_{r-1}(u_j^{(r)}+\ii/2)}
  \frac{Q_{r+1}(u_j^{(r)}-\ii/2)}
       {Q_{r+1}(u_j^{(r)}+\ii/2)} .
\]
The factor involving \(Q_r\) contains the root \(u_j^{(r)}\) itself:
\[
  -\frac{Q_r(u_j^{(r)}+\ii)}{Q_r(u_j^{(r)}-\ii)}
  =
  \prod_{\substack{k=1\\k\ne j}}^{M_r}
  \frac{u_j^{(r)}-u_k^{(r)}+\ii}
       {u_j^{(r)}-u_k^{(r)}-\ii}.
\]
Combining this same-node factor with the two adjacent-node factors is exactly
the Cartan-matrix product in \eqref{eq:twisted-su-n-dva-bethe}.  For the
homogeneous untwisted chain,
\[
  \frac{\phi_1(u_j^{(1)}-\ii/2)}
       {\phi_2(u_j^{(1)}-\ii/2)}
  =
  \left(\frac{u_j^{(1)}+\ii/2}{u_j^{(1)}-\ii/2}\right)^L,
  \qquad
  \frac{\phi_r}{\phi_{r+1}}=1\quad(r\ge2),
\]
which gives \eqref{eq:su-n-nested-bethe-cartan}.

Formula \eqref{eq:su-n-dressed-vacuum-eigenvalue} is often the cleanest way
to compare the algebraic Bethe ansatz, functional \(TQ/QQ\) systems, and
separation of variables.  The finite-dimensional transfer matrix is
polynomial after all pole residues vanish; polynomiality, not a separate
appeal to a formal Bethe equation, is the algebraic reason the spectrum is
well-defined.

\section{Worked SU(3) Configuration}

For \(SU(3)\) there are first-level roots \(u_j\) and second-level roots
\(v_\alpha\).  The equations are
\begin{align}
  \left(\frac{u_j+\ii/2}{u_j-\ii/2}\right)^L
  &=
  \prod_{\substack{k=1\\ k\neq j}}^{M_1}
  \frac{u_j-u_k+\ii}{u_j-u_k-\ii}
  \prod_{\alpha=1}^{M_2}
  \frac{u_j-v_\alpha-\ii/2}{u_j-v_\alpha+\ii/2},
  \label{eq:su3-first-level}\\
  1
  &=
  \prod_{\substack{\beta=1\\ \beta\neq\alpha}}^{M_2}
  \frac{v_\alpha-v_\beta+\ii}{v_\alpha-v_\beta-\ii}
  \prod_{j=1}^{M_1}
  \frac{v_\alpha-u_j-\ii/2}{v_\alpha-u_j+\ii/2}.
  \label{eq:su3-second-level}
\end{align}
Set \(L=6\), \(M_1=2\), \(M_2=1\),
\[
  u_1=-\frac{\sqrt3}{2},\qquad
  u_2=\frac{\sqrt3}{2},\qquad
  v_1=0 .
\]
Let
\[
  z=\frac{u_2+\ii/2}{u_2-\ii/2}=\ee^{\ii\pi/3}.
\]
Then \(z^6=1\).  The second-level equation is
\[
  \frac{\sqrt3/2-\ii/2}{\sqrt3/2+\ii/2}
  \frac{-\sqrt3/2-\ii/2}{-\sqrt3/2+\ii/2}
  =
  z^{-1}z=1 .
\]
For the first root \(u_2\), the right side of
\eqref{eq:su3-first-level} is
\[
  \frac{2u_2+\ii}{2u_2-\ii}
  \frac{u_2-\ii/2}{u_2+\ii/2}
  =
  z\,z^{-1}=1,
\]
and the left side is \(z^6=1\).  The equation for \(u_1\) is the inverse of
the same equality.  The energy is \(E=2\).  This example is deliberately
small: every phase can be checked exactly, and the second-level root is
essential even though it carries no direct energy.

\section{Trigonometric Nesting}

The relativistic sine-Gordon and affine-Toda families use trigonometric
matrix parts rather than the rational \(R(u)=u+\ii P\).  A convenient
\(\mathfrak{sl}_N\) trigonometric normalization is
\[
\begin{aligned}
  R_\eta(u)
  &=
  \sinh(u+\eta)\sum_a E_{aa}\otimes E_{aa}
  +
  \sinh u\sum_{a\ne b}E_{aa}\otimes E_{bb}
\\
  &\quad
  +
  \sinh\eta
  \sum_{a\ne b}\ee^{\operatorname{sgn}(a-b)u}
  E_{ab}\otimes E_{ba}.
\end{aligned}
\]
It solves the Yang--Baxter equation and degenerates to the rational matrix
after \(u,\eta\to0\) with \(u/\eta\) fixed, up to scalar normalization.  The
nested pole-cancellation equations become
\begin{equation}
  \left(
    \frac{\sinh(u_j^{(r)}+\frac{\eta}{2}\delta_{r1})}
         {\sinh(u_j^{(r)}-\frac{\eta}{2}\delta_{r1})}
  \right)^L
  =
  \prod_{s=1}^{N-1}
  \prod_{\substack{k=1\\(s,k)\ne(r,j)}}^{M_s}
  \frac{\sinh(u_j^{(r)}-u_k^{(s)}+\frac{\eta}{2}A_{rs})}
       {\sinh(u_j^{(r)}-u_k^{(s)}-\frac{\eta}{2}A_{rs})}.
  \label{eq:trigonometric-nested-bethe-cartan}
\end{equation}
The proof is the same recursive RTT proof as
Theorem~\ref{thm:recursive-nested-bethe-origin}, with rational functions
replaced by hyperbolic functions.  The scalar factor multiplying
\(R_\eta\) is not fixed by the Yang--Baxter equation; in a relativistic QFT it
is fixed separately by unitarity, crossing, bound-state residues, and
normalization of asymptotic states.

\section{Relativistic SU(N)-Symmetric Scattering}

For a relativistic integrable QFT whose particles transform in a fundamental
\(SU(N)\) multiplet, the two-body \(S\)-matrix often factorizes as
\[
  S(\theta)=S_0(\theta)\,\widehat R(\theta),
\]
where \(S_0\) is a scalar meromorphic factor and \(\widehat R\) is an
\(SU(N)\)-invariant rational or trigonometric matrix normalized to satisfy
unitarity and crossing in the physical rapidity variable \(\theta\).  The
finite-volume quantization separates into a scalar phase and a nested
transfer-matrix eigenvalue:
\begin{equation}
  mL\sinh\theta_j
  -\ii\sum_{k\neq j}\log S_0(\theta_j-\theta_k)
  -\ii\log \tau_j(\theta_1,\ldots,\theta_n;\{u^{(r)}\})
  =
  2\pi I_j .
  \label{eq:relativistic-nested-bethe-yang}
\end{equation}
The auxiliary roots \(\{u^{(r)}\}\) obey rational or trigonometric equations
of the form \eqref{eq:su-n-nested-bethe-cartan} or
\eqref{eq:trigonometric-nested-bethe-cartan}, with the map from physical
rapidity \(\theta\) to auxiliary spectral parameter determined by
\(\widehat R\).  Changing this map by a meromorphic reparametrization changes
the displayed coordinate form but not the transfer-matrix eigenvalue.

\begin{controlledapproximation}[Large-volume Bethe--Yang regime]
\label{ca:matrix-bethe-yang-large-volume}
Equation~\eqref{eq:relativistic-nested-bethe-yang} is a large-\(L\)
quantization condition for stable particles whose separations are much larger
than the interaction range set by the mass gap.  Corrections are
exponentially small in \(mL\) for a massive theory without massless channels,
provided no bound-state pole is pinched by the finite-volume contour.  Power
corrections and wrapping corrections require the TBA or mirror-channel
formalism of later chapters.
\end{controlledapproximation}

\section{Nested Strings and Thermodynamic Input}

In a thermodynamic limit the roots of nested equations often organize into
strings.  In the rational normalization, an \(m\)-string at node \(r\) with
real center \(x\) has the asymptotic form
\begin{equation}
  u_{\ell}^{(r,m)}
  =
  x+\frac{\ii}{2}(m+1-2\ell)+\varepsilon_\ell(L),
  \qquad \ell=1,\ldots,m,
  \label{eq:nested-string-pattern}
\end{equation}
with \(\varepsilon_\ell(L)\to0\) in the regime where the string hypothesis is
valid.  The string center becomes the variable in the thermodynamic density
equations; the string length \(m\) and Dynkin node \(r\) become species
labels.

\begin{hypothesis}[Nested string hypothesis]
\label{ca:nested-string-hypothesis}
The replacement of finite-volume roots by exact string species is not a
formal consequence of the Bethe equations alone.  It is a model-dependent
thermodynamic hypothesis requiring control of singular strings, deviations
\(\varepsilon_\ell(L)\), possible holes, and the order of limits.  When used
below, the input is: for the thermodynamic states under consideration, all
roots not on real physical rapidity lines can be partitioned into families of
the form \eqref{eq:nested-string-pattern}, with deviations whose contribution
to intensive free energy vanishes in the limit.  This is the precise
assumption that converts the finite nested Bethe equations into the
multi-species TBA equations of
Chapter~\ref{ch:integrable-nested-tba-tq-separation-variables}.
\end{hypothesis}

\section{Principal Chiral Model and Gross--Neveu Families}

The \(SU(N)\) principal chiral model has global
\(SU(N)_L\times SU(N)_R\) symmetry.  Its massive particles carry left and
right internal indices, so the matrix part of its scattering problem contains
two nested \(SU(N)\) systems, one for each chiral factor.  Gross--Neveu
families have a similar nested structure, with the details controlled by the
vector, spinor, or fundamental representation carried by the particles.  The
Bethe equations therefore contain physical rapidities together with auxiliary
roots attached to the Dynkin diagram of the relevant symmetry algebra.

The important conceptual point is that auxiliary roots are not additional
particles in the Hilbert space.  They parametrize the internal wavefunction
of a multiparticle state.  The energy and momentum are functions of the
physical rapidities; the auxiliary roots enter by selecting the
transfer-matrix eigenvalue that appears in the quantization condition.

\section{Completeness and Its Precise Scope}

\begin{quotedtheorem}[Generic finite-chain completeness]
\label{qthm:generic-nested-bethe-completeness}
\emph{Proof references.}\footnote{The algebraic nested Bethe vectors and
higher transfer-matrix eigenvalues for rational \(\mathfrak{gl}_N\)
XXX-type models are developed in Mukhin--Tarasov--Varchenko
\cite{MTV2006BetheEigenvectors}.  Simple-spectrum and completeness results
for XXX Bethe algebras are proved in the Mukhin--Tarasov--Varchenko and
Tarasov works cited in
\cite{MTV2009XXXSimpleSpectrum,Tarasov2018Completeness}.  The formulation
used here is the generic finite-chain theorem-boundary form: generic twist
and inhomogeneity are hypotheses, not decorative parameters.}
For the inhomogeneous rational \(SU(N)\) fundamental chain with a generic
diagonal twist whose eigenvalue ratios are not roots of unity, and with
generic inhomogeneities, the nested Bethe equations have a set of admissible
solutions whose Bethe vectors form a basis of \((\C^N)^{\otimes L}\).  The
transfer-matrix spectrum is simple on a Zariski-open set of the twist and
inhomogeneity parameters.
\end{quotedtheorem}

\paragraph{Mechanism and singular-limit boundary.}
The quoted theorem is a finite-dimensional algebraic theorem about the
commutative Bethe algebra generated by transfer matrices.  The generic
parameters serve two purposes.  First, they separate transfer-matrix
eigenvalues: on the Zariski-open parameter set the joint spectrum is simple,
so an eigenvector is determined by its eigenvalue data.  Second, the
inhomogeneities and twist remove the singular Bethe-root configurations that
make the coordinate Bethe equations incomplete at special points.  The proof
strategy can be summarized as follows.  The transfer matrices generate a
commutative algebra whose joint eigenvalues are encoded by a rational
difference operator or, equivalently, by Wronskian/Casoratian polynomial data.
For generic parameters the relevant algebraic variety is reduced and has the
dimension predicted by the representation space.  Admissible Bethe solutions
map to this variety, the Bethe vectors attached to distinct solutions have
distinct joint eigenvalues, and dimension counting gives a basis.  None of
these steps is a thermodynamic statement; every step takes place at fixed
finite chain length \(L\).

The theorem is a finite-dimensional algebraic statement.  It does not by
itself prove completeness for the periodic homogeneous chain, because taking
the twist to the identity and the inhomogeneities to equal values can create
singular roots and degeneracies.  It also does not prove that a continuum QFT
finite-volume Hamiltonian is approximated by matrix Bethe--Yang equations.
For continuum use we therefore separate three assertions:
\[
\begin{array}{ll}
\text{algebraic completeness} & \text{basis theorem for a finite transfer matrix,}\\
\text{Bethe--Yang control} & \text{large-volume approximation to QFT levels,}\\
\text{thermodynamic control} & \text{existence of the density/string limit.}
\end{array}
\]
Each assertion has different hypotheses.  The monograph uses the generic
finite-chain theorem as an algebraic guide, but it states explicitly whenever
periodic homogeneous completeness, continuum spectral control, or the string
hypothesis is being assumed rather than proved.

\section{Interface with Later Volumes}

The relativistic integrable-QFT use of nested TBA differs from planar
gauge-theory integrability.  In the present volume, physical particles have
relativistic dispersion \(E=m\cosh\theta\) and finite-size effects are
organized by the mirror channel of a two-dimensional QFT.  In the planar
\(\mathcal N=4\) SYM volume, magnons have a non-relativistic dispersion on a
spin chain, wrapping effects arise from trace cyclicity and finite operator
length, and the quantum spectral curve replaces the finite-rank Baxter
polynomial by a system of analytic \(Q\)-functions.  The algebraic bridge is
the RTT, nested monodromy, \(TQ\), \(QQ\), and Hirota logic developed here
and in Chapter~\ref{ch:integrable-nested-tba-tq-separation-variables}; the
physical interpretation is different.
