\chapter{Bridges To Nonintegrable Two-Dimensional QFT And CFT}
\label{ch:integrable-bridges-nonintegrable-two-dimensional-qft}

\ConventionsMixedCFT

Exact integrability supplies unusually explicit coordinates on a local QFT:
asymptotic particles, factorized scattering, form factors, finite-volume
Bethe--Yang spectra, and TBA vacuum energies.  A nearby nonintegrable theory
is studied by deforming these data through a declared regulator and local
operator chart.  The bridge developed in this chapter is local and
finite-volume first; infinite-volume resonance poles and decay widths are
defined as limits of finite-volume matrix elements.

\section{Integrability-Breaking Datum}

Let \(\mathcal Q_0\) be a massive integrable two-dimensional QFT with Hilbert
space \(\Hilb_0\), Hamiltonian \(H_0\), local operator algebra
\(\mathcal A_0\), and commuting conserved charges
\[
  Q_s,\qquad s\in\mathcal S\subset\mathbb Z .
\]
The charge \(Q_s\) acts diagonally on an asymptotic multiparticle state
\[
  \alpha=(a_1,\theta_1;\ldots;a_n,\theta_n)
\]
with eigenvalue
\[
  q_s(\alpha)=\sum_{j=1}^n q_s^{(a_j)}(\theta_j),
\]
where \(q_s^{(a)}(\theta)\) is the one-particle charge eigenvalue.  For
ordinary energy and momentum,
\[
  q_1^{(a)}(\theta)=m_a\cosh\theta,\qquad
  \widetilde q_1^{(a)}(\theta)=m_a\sinh\theta .
\]

\begin{definition}[Regulated integrability-breaking deformation]
\label{def:regulated-integrability-breaking-deformation}
A regulated deformation of \(\mathcal Q_0\) is a family of finite-volume
Hamiltonians
\[
  H(\epsilon,L,\Lambda)
  =
  H_0(L,\Lambda)
  +
  \epsilon V_b(L,\Lambda)
  +
  \sum_A c_A(\epsilon,\Lambda)\,V_A(L,\Lambda),
\]
where
\[
  V_b(L,\Lambda)=\int_0^L dx\,\mathcal O_b(x)_\Lambda,
  \qquad
  V_A(L,\Lambda)=\int_0^L dx\,\mathcal O_A(x)_\Lambda .
\]
Here \(L\) is the spatial circumference, \(\Lambda\) is a UV cutoff,
\(\mathcal O_b\) is the declared perturbing local operator, and the finite
set of local coordinates \(c_A\) is fixed by the chosen regularization and
renormalization condition.  The continuum deformation is the limit of this
family in matrix elements or Schwinger functions after these local
coordinates have been specified.
\end{definition}

The deformation parameter \(\epsilon\) has engineering dimension
\[
  [\epsilon]=2-\Delta_b
\]
when \(\mathcal O_b\) has scaling dimension \(\Delta_b\) at a conformal
short-distance fixed point.  Away from a CFT, the same symbol denotes the
coordinate multiplying the integrated local operator in the regulated
Hamiltonian.

\section{Broken Conservation Laws}

Let \(V_b=\int dx\,\mathcal O_b(x)\) in infinite volume, with a finite-volume
regulator understood.  At \(\epsilon=0\),
\[
  [H_0,Q_s]=0 .
\]
The deformed time derivative of \(Q_s\) is
\[
  \frac{dQ_s}{dt}
  =
  \ii[H(\epsilon),Q_s]
  =
  \ii\epsilon[V_b,Q_s]+O(\epsilon^2).
\]
If \(Q_s=\int dx\,j_s^0(x)\), this is the integrated form of
\[
  \partial_\mu j_s^\mu(x)=\epsilon\,\mathcal B_s(x)+O(\epsilon^2),
\]
where \(\mathcal B_s\) is the local operator representing the charge-breaking
density in the chosen operator chart.

\begin{proposition}[Broken-charge matrix element ledger]
\label{prop:broken-charge-matrix-ledger}
Let \(\alpha\) and \(\beta\) be finite-particle scattering states of the
integrable theory on which \(Q_s\) has eigenvalues \(q_s(\alpha)\) and
\(q_s(\beta)\).  Then the first-order broken-charge matrix element obeys
\[
  \langle\beta|\ii[V_b,Q_s]|\alpha\rangle
  =
  \ii\bigl(q_s(\alpha)-q_s(\beta)\bigr)
  \langle\beta|V_b|\alpha\rangle .
\]
\end{proposition}

\begin{proof}
Since \(Q_s|\alpha\rangle=q_s(\alpha)|\alpha\rangle\) and
\(\langle\beta|Q_s=q_s(\beta)\langle\beta|\),
\[
  \langle\beta|V_bQ_s|\alpha\rangle
  =
  q_s(\alpha)\langle\beta|V_b|\alpha\rangle,
  \qquad
  \langle\beta|Q_sV_b|\alpha\rangle
  =
  q_s(\beta)\langle\beta|V_b|\alpha\rangle .
\]
Subtracting the two identities and multiplying by \(\ii\) gives the formula.
\end{proof}

The equation is useful because the integrable charges give many independent
kinematic tests.  A matrix element that changes any higher-spin charge
eigenvalue is a direct channel through which the deformation breaks
factorized scattering.  Energy and momentum remain conserved when the
deformed Hamiltonian is translation invariant; higher charges typically do
not.

\section{Form-Factor Perturbation Theory}

Let \(\ket{A_a(\theta)}\) denote a one-particle state of species \(a\) and
rapidity \(\theta\) in the integrable theory.  The Lorentzian normalization
used in this volume is
\[
  \langle A_a(\theta')|A_b(\theta)\rangle
  =
  2\pi\,\delta_{ab}\delta(\theta'-\theta).
\]
For a scalar perturbing operator \(\mathcal O_b\), define the connected
diagonal crossed form factor by
\[
  F_{a\bar a}^{b,\mathrm{conn}}(\theta+\ii\pi,\theta)
  =
  {}_{\mathrm{out}}\!\langle A_a(\theta)|
  \mathcal O_b(0)
  |A_a(\theta)\rangle_{\mathrm{in,conn}} .
\]
Lorentz invariance makes this connected diagonal matrix element independent
of \(\theta\) for a scalar operator; denote it by \(F_a^b\).

\begin{proposition}[First-order mass shift from a diagonal form factor]
\label{prop:ffpt-first-order-mass-shift}
Assume that the one-particle state remains an isolated stable particle under
the deformation, and that the diagonal connected form factor \(F_a^b\) is
finite in the declared crossing prescription.  For
\[
  H=H_0+\epsilon\int dx\,\mathcal O_b(x)+O(\epsilon^2),
\]
the first-order shift of the squared mass is
\[
  \delta m_a^2=2\epsilon F_a^b .
\]
\end{proposition}

\begin{proof}
Place the theory on a circle of circumference \(L\).  The one-particle
Bethe--Yang density is
\[
  \rho_1(\theta)=m_aL\cosh\theta .
\]
The finite-volume diagonal form-factor formula of
Chapter~\ref{ch:integrable-finite-volume-form-factors} gives, after the
vacuum term is separated into the bulk energy coordinate,
\[
  {}_L\langle A_a(\theta)|\mathcal O_b(0)|A_a(\theta)\rangle_L^{\mathrm{conn}}
  =
  \frac{F_a^b}{m_aL\cosh\theta}
  +O(e^{-\mu L}).
\]
Thus the connected first-order energy shift is
\[
  \delta E_a(\theta)
  =
  \epsilon L
  {}_L\langle A_a(\theta)|\mathcal O_b(0)|A_a(\theta)\rangle_L^{\mathrm{conn}}
  =
  \epsilon\frac{F_a^b}{m_a\cosh\theta}
  +O(e^{-\mu L}).
\]
If the shifted dispersion remains relativistic,
\[
  E_a(\theta)^2-p_a(\theta)^2=m_a^2+\delta m_a^2+O(\epsilon^2),
\]
then at fixed momentum
\[
  \delta E_a(\theta)
  =
  \frac{\delta m_a^2}{2E_a(\theta)}
  =
  \frac{\delta m_a^2}{2m_a\cosh\theta}.
\]
Comparing the two expressions gives \(\delta m_a^2=2\epsilon F_a^b\).
\end{proof}

\begin{definition}[Form-factor perturbation datum]
\label{def:form-factor-perturbation-datum}
A first-order form-factor perturbation datum consists of:
\begin{enumerate}
\item the perturbing operator normalization in the local operator chart;
\item the crossing prescription for diagonal form factors and kinematic-pole
subtractions;
\item the vacuum-energy subtraction convention;
\item the finite-volume normalization used to take the infinite-volume limit;
\item the stability hypothesis specifying which isolated particles remain
asymptotic states.
\end{enumerate}
\end{definition}

The stability hypothesis is part of the datum because an integrable particle
may become a resonance or may be confined after the deformation.  The finite
diagonal form factor then supplies a mass counterterm or a pole displacement
only after the relevant analytic continuation has been declared.

\section{Semi-Locality and Confinement Diagnostics}

Let \(\mathcal O_b\) be semi-local with respect to particle species \(a\).
The semi-locality phase \(\omega_{b,a}\in U(1)\) is the monodromy acquired
when the particle-creating field for \(a\) is moved once around
\(\mathcal O_b\).  The kinematic pole equation of the form-factor chapter
then gives the two-particle diagonal pole
\[
  -\ii\,\operatorname*{Res}_{\theta'=\theta+\ii\pi}
  F_{\bar a a}^{b}(\theta',\theta)
  =
  \left(1-\omega_{b,a}\right)\langle\mathcal O_b\rangle .
\]

\begin{proposition}[Semi-local pole obstruction to a particle mass shift]
\label{prop:semilocal-pole-confinement-diagnostic}
If \(\omega_{b,a}\ne1\) and
\(\langle\mathcal O_b\rangle\ne0\), the first-order diagonal form factor
entering Proposition~\ref{prop:ffpt-first-order-mass-shift} has a kinematic
pole at the required crossing point.  The perturbative one-particle mass
shift is therefore replaced by a confinement or resonance problem in the
deformed theory.
\end{proposition}

\begin{proof}
The displayed residue is nonzero under the stated hypotheses.  Hence the
crossed two-particle form factor has a pole at
\(\theta'=\theta+\ii\pi\), which is exactly the diagonal point used to define
the connected one-particle matrix element.  The finite value \(F_a^b\) in
Proposition~\ref{prop:ffpt-first-order-mass-shift} is absent.  Thus the
assumption that the original one-particle state admits an ordinary first-order
mass shift fails.  The deformed spectral problem must be formulated with the
appropriate long-distance force, resonance pole, or finite-volume mixing
included.
\end{proof}

\begin{example}[Ising magnetic perturbation and kink confinement]
\label{ex:ising-magnetic-kink-confinement}
In the ordered phase of the thermally perturbed Ising model there are two
vacua \(|+\rangle\) and \(|-\rangle\) with
\[
  \langle+\!|\sigma|+\rangle=v,\qquad
  \langle-\!|\sigma|-\rangle=-v .
\]
Perturb the Hamiltonian by
\[
  \delta H=-h\int dx\,\sigma(x).
\]
The two vacuum energy densities shift by \(-hv\) and \(+hv\).  A kink--antikink
pair separated by distance \(R\) encloses a segment of the higher-energy
vacuum.  The leading long-distance potential is therefore
\[
  V_{\mathrm{conf}}(R)=2|hv|\,R .
\]
For two heavy kinks of mass \(m\), the leading nonrelativistic meson equation
for the relative coordinate is
\[
  \left[
  -\frac1m\frac{d^2}{dR^2}
  +2|hv|\,|R|
  \right]\psi(R)
  =
  (M-2m)\psi(R).
\]
This equation is an infrared consequence of the vacuum energy splitting.  The
operator-level warning sign is the semi-local kinematic pole of the spin-field
form factor with respect to the kink.
\end{example}

\section{Finite-Volume Transition Rates}

Let
\[
  V_b(L)=\int_0^L dx\,\mathcal O_b(x)
\]
and let \(\ket{A,L}\) be an eigenstate of \(H_0(L)\).  Let
\(\ket{n,L}\) be finite-volume eigenstates with energies \(E_n(L)\).  First
order time-dependent perturbation theory gives
\[
  P_{A\to n}(T,L)
  =
  \epsilon^2
  \left|
    \int_0^T dt\,e^{\ii(E_n-E_A)t}
    \langle n,L|V_b(L)|A,L\rangle
  \right|^2
  +
  O(\epsilon^3).
\]
The elementary distributional limit is
\[
  \frac1T
  \left|\int_0^Tdt\,e^{\ii\Delta E t}\right|^2
  =
  \frac{4\sin^2(\Delta E T/2)}{T(\Delta E)^2}
  \longrightarrow
  2\pi\delta(\Delta E).
\]

\begin{proposition}[Finite-volume golden-rule limit]
\label{prop:finite-volume-golden-rule-limit}
Assume that the \(L\to\infty\) limit of the finite-volume spectral sum exists
after the Bethe--Yang density of states is inserted and that the relevant
crossed form factors are regular on the energy-conserving contour.  For a
channel
\[
  \beta=(b_1,\theta_1;\ldots;b_n,\theta_n)
\]
use the momentum-conserving rapidity measure
\[
  d\Phi_\beta
  =
  2\pi\delta
  \!\left(P_A-\sum_{j=1}^n m_{b_j}\sinh\theta_j\right)
  \prod_{j=1}^n\frac{d\theta_j}{2\pi}.
\]
Then the leading inclusive decay rate out of \(A\) is
\[
  \Gamma_A
  =
  2\pi\epsilon^2
  \sum_\beta
  \int d\Phi_\beta\,
  \delta\!\left(E_A-\sum_{j=1}^n m_{b_j}\cosh\theta_j\right)
  \left|
    \langle\beta|\mathcal O_b(0)|A\rangle_{\mathrm{conn}}
  \right|^2 ,
\]
\end{proposition}

\begin{proof}
At finite \(L\), the transition probability per time is the finite sum over
normalizable states
\[
  \frac{P_A(T,L)}{T}
  =
  \epsilon^2
  \sum_n
  \frac{4\sin^2((E_n-E_A)T/2)}{T(E_n-E_A)^2}
  |\langle n,L|V_b(L)|A,L\rangle|^2
  +O(\epsilon^3).
\]
First take \(T\to\infty\) in the distributional sense against smooth spectral
test functions, obtaining \(2\pi\delta(E_n-E_A)\).  Then use the
Bethe--Yang state-counting measure of
Chapter~\ref{ch:integrable-finite-volume-form-factors} to replace the sum by
rapidity integrals.  The spatial integral in \(V_b(L)\) gives finite-volume
momentum conservation; in the infinite-volume limit this gives the delta
function contained in \(d\Phi_\beta\).  The finite-volume matrix-element
normalization converges to the connected crossed form factor by the same
Gaudin-density cancellation.  This proves the displayed inclusive rate.
\end{proof}

\section{Two-Particle Phase Space In Two Dimensions}

The finite-volume formula becomes especially sharp for the first allowed
decay channel.  Let the unstable particle \(A\) have mass \(M_A\), and study
the contribution of two stable particles \(r,s\) of masses \(m_r,m_s\).  In
the rest frame of \(A\), energy-momentum conservation has a real solution
precisely when
\[
  M_A>m_r+m_s .
\]
Define the Kallen polynomial
\[
  \lambda(x,y,z)=x^2+y^2+z^2-2xy-2xz-2yz
\]
and the relative momentum
\[
  p_*=
  \frac{\sqrt{\lambda(M_A^2,m_r^2,m_s^2)}}{2M_A}.
\]
The two on-shell energies at the decay point are
\[
  E_r^*=\frac{M_A^2+m_r^2-m_s^2}{2M_A},
  \qquad
  E_s^*=\frac{M_A^2+m_s^2-m_r^2}{2M_A}.
\]
For each orientation \(\sigma=\pm1\), let
\[
  m_r\sinh\theta_r^\sigma=\sigma p_*,
  \qquad
  m_s\sinh\theta_s^\sigma=-\sigma p_* .
\]

\begin{proposition}[Two-particle decay width in \(1+1\) dimensions]
\label{prop:two-particle-decay-width-two-d}
Use the state normalization and golden-rule convention of
Proposition~\ref{prop:finite-volume-golden-rule-limit}.  Suppose \(A\) is at
rest, \(M_A>m_r+m_s\), and the crossed connected form factor
\[
  \mathcal F_{rs;A}(\theta_r,\theta_s)
  =
  \langle A_r(\theta_r)A_s(\theta_s)|
  \mathcal O_b(0)|A\rangle_{\mathrm{conn}}
\]
has a finite boundary value at the two energy-conserving rapidity pairs
\((\theta_r^\sigma,\theta_s^\sigma)\).  If the channel sum is over labelled
ordered particles, the \(rs\) contribution to the leading width is
\[
  \Gamma_{A\to rs}
  =
  \frac{\epsilon^2}{M_Ap_*}
  \sum_{\sigma=\pm1}
  \left|
    \mathcal F_{rs;A}(\theta_r^\sigma,\theta_s^\sigma)
  \right|^2 .
\]
If instead the channel is counted as an unordered physical channel, the
right-hand side must be divided by \(2\) when \(r=s\), and is unchanged when
\(r\ne s\).
\end{proposition}

\begin{proof}
Start from Proposition~\ref{prop:finite-volume-golden-rule-limit} with
\(P_A=0\) and with a two-particle labelled channel:
\[
  \Gamma_{A\to rs}
  =
  2\pi\epsilon^2
  \int
  \frac{d\theta_r}{2\pi}
  \frac{d\theta_s}{2\pi}
  2\pi\delta(p_r+p_s)
  \delta(M_A-E_r-E_s)
  |\mathcal F_{rs;A}(\theta_r,\theta_s)|^2 ,
\]
where
\[
  p_i=m_i\sinh\theta_i,\qquad E_i=m_i\cosh\theta_i .
\]
Change variables from rapidities to momenta.  Since \(d\theta_i=dp_i/E_i\),
the phase-space integral without the exterior factor \(2\pi\epsilon^2\) is
\[
  \int_{\R^2}
  \frac{dp_r\,dp_s}{2\pi E_rE_s}\,
  \delta(p_r+p_s)\,
  \delta(M_A-E_r-E_s)\,
  |\mathcal F_{rs;A}|^2 .
\]
The momentum delta function sets \(p_r=p\) and \(p_s=-p\).  Therefore
\[
  E_r(p)=\sqrt{m_r^2+p^2},\qquad
  E_s(p)=\sqrt{m_s^2+p^2}.
\]
The equation \(E_r(p)+E_s(p)=M_A\) has the two roots
\(p=\sigma p_*\), \(\sigma=\pm1\), because solving it gives
\[
  E_r^*=\frac{M_A^2+m_r^2-m_s^2}{2M_A},
  \qquad
  (E_r^*)^2-m_r^2
  =
  \frac{\lambda(M_A^2,m_r^2,m_s^2)}{4M_A^2}.
\]
At either root,
\[
  \left|
  \frac{d}{dp}(E_r(p)+E_s(p))
  \right|
  =
  |p_*|\left(\frac1{E_r^*}+\frac1{E_s^*}\right)
  =
  \frac{|p_*|M_A}{E_r^*E_s^*}.
\]
Thus each root contributes
\[
  \frac{1}{2\pi E_r^*E_s^*}\,
  \frac{E_r^*E_s^*}{|p_*|M_A}
  |\mathcal F_{rs;A}(\theta_r^\sigma,\theta_s^\sigma)|^2 .
\]
Multiplication by the exterior \(2\pi\epsilon^2\) gives the displayed
formula.  If \(r=s\) and the physical channel is unlabelled, the two labels
of the identical particles describe the same basis of final states twice;
dividing by \(2\) implements the chosen unordered channel convention.
\end{proof}

The factor \(1/p_*\) is the one-dimensional threshold singularity of the
two-particle density of states.  It is not by itself a physical divergence in
an observable statement: at threshold the perturbative decay problem must be
examined with the form-factor boundary value, finite-volume level spacing,
and possible near-threshold bound or virtual-state poles kept in the same
scaling limit.

\section{Truncated Conformal Space as Regulator}

Let \(\mathcal C\) be a two-dimensional CFT on a circle of circumference
\(L\).  The CFT Hamiltonian is
\[
  H_{\mathrm{CFT}}
  =
  \frac{2\pi}{L}
  \left(L_0+\bar L_0-\frac{c}{12}\right).
\]
Perturbing by relevant scalar operators \(\mathcal O_i\) gives
\[
  H
  =
  H_{\mathrm{CFT}}
  +
  \sum_i \lambda_i
  \left(\frac{2\pi}{L}\right)^{\Delta_i-1}
  \int_0^{2\pi}d\theta\,\mathcal O_i(\theta).
\]
Equivalently,
\[
  \frac{L}{2\pi}H
  =
  L_0+\bar L_0-\frac{c}{12}
  +
  \sum_i g_i(L)\int_0^{2\pi}d\theta\,\mathcal O_i(\theta),
  \qquad
  g_i(L)=\lambda_i\left(\frac{L}{2\pi}\right)^{2-\Delta_i}.
\]

\begin{definition}[TCSA regulator datum]
\label{def:tcsa-regulator-datum}
A truncated conformal space regulator consists of:
\begin{enumerate}
\item a CFT Hilbert-space sector on the circle;
\item a cutoff subspace \(P_N\Hilb_{\mathcal C}\) defined by
\(L_0+\bar L_0\le N\) or by a stated energy cutoff;
\item the matrix elements of all perturbing operators in this CFT basis;
\item local counterterm coordinates determined by the discarded high-energy
states;
\item an order of limits for \(N\), \(L\), and the renormalized couplings.
\end{enumerate}
\end{definition}

The counterterm scaling follows from the OPE.  Suppose
\[
  \mathcal O_i(x)\mathcal O_j(0)
  \sim
  \sum_k
  C_{ij}{}^k\,|x|^{\Delta_k-\Delta_i-\Delta_j}\mathcal O_k(0).
\]
The short-distance part of the second-order perturbation contains
\[
  \int_{|x|<a}d^2x\,|x|^{\Delta_k-\Delta_i-\Delta_j}
  =
  2\pi
  \frac{a^{2+\Delta_k-\Delta_i-\Delta_j}}
       {2+\Delta_k-\Delta_i-\Delta_j}
\]
when the denominator is nonzero, and a logarithm when it vanishes.  With
\(a\sim N^{-1}\), the cutoff dependence of the local coordinate multiplying
\(\mathcal O_k\) scales as
\[
  \delta g_k(N)
  \propto
  \lambda_i\lambda_j\,
  N^{\Delta_i+\Delta_j-\Delta_k-2}
\]
up to scheme-dependent constants and logarithmic replacement at exponent
zero.

\begin{proposition}[OPE origin of TCSA counterterm powers]
\label{prop:tcsa-counterterm-power}
In a TCSA regulator whose high-energy cutoff corresponds to a short-distance
resolution \(a\sim N^{-1}\), the contribution of the OPE channel
\(\mathcal O_i\mathcal O_j\to\mathcal O_k\) to a local second-order
counterterm has cutoff power
\[
  N^{\Delta_i+\Delta_j-\Delta_k-2}.
\]
If \(\Delta_i+\Delta_j-\Delta_k-2=0\), the power is replaced by \(\log N\).
\end{proposition}

\begin{proof}
Second-order perturbation theory integrates the product
\(\mathcal O_i(x)\mathcal O_j(0)\) over small Euclidean separations.  Replacing
the product by its OPE channel gives the radial integral displayed above.  If
the exponent
\[
  2+\Delta_k-\Delta_i-\Delta_j
\]
is nonzero, substituting \(a=N^{-1}\) gives
\[
  a^{2+\Delta_k-\Delta_i-\Delta_j}
  =
  N^{\Delta_i+\Delta_j-\Delta_k-2}.
\]
If the exponent vanishes, the radial integral is
\(\int_a dr\,r^{-1}\), giving \(\log N\).  The coefficient is scheme
dependent because the precise relation between \(a\) and \(N\), and the
definition of the projection \(P_N\), are part of the regulator.
\end{proof}

\section{Comparison Objects}

A bridge between integrable and nonintegrable two-dimensional QFT compares
declared observables:
\[
\begin{gathered}
  \text{finite-volume spectra},\qquad
  \text{local-operator matrix elements},\\
  \text{resonance pole positions and residues},\qquad
  \text{thermal or hydrodynamic late-time data}.
\end{gathered}
\]
An integrable point controls nearby data through form-factor perturbation
theory in \(\epsilon\).  A CFT point controls nearby data through relevant
deformations and Hamiltonian truncation.  The two coordinate systems meet
when the finite-volume regulator, local counterterms, operator normalization,
and infinite-volume continuation are matched.

\begin{openproblem}[Nonintegrable deformation]
\label{op:controlled-nonintegrable-deformation}
Construct a nonintegrable two-dimensional QFT near an integrable point with a
finite-volume regulator, local counterterm coordinates, form-factor
perturbation theory, resonance pole analysis, and a proof of convergence to a
local continuum QFT in a specified scaling domain.
\end{openproblem}
