\chapter{Two-Dimensional Scattering Analyticity and Bootstrap Data}
\label{ch:integrable-scattering-analyticity-bootstrap}

\ConventionsLorentzian

The factorized scattering chapter fixed the algebraic data of an integrable
massive theory.  This chapter adds the analytic data in the rapidity plane.
The object studied here is an on-shell two-body scattering function together
with analytic hypotheses.  A local QFT realizing that function is a further
object; its construction requires Hilbert-space, locality, and domain data.

\section{Rapidity Plane and Physical Strip}

For equal masses \(m\), the two-particle Mandelstam invariant is
\[
  s(\theta)=2m^2(1+\cosh\theta)
  =4m^2\cosh^2\!\left(\frac{\theta}{2}\right).
\]
The physical two-particle scattering line is \(\theta\in\R\).  The crossed
channel is reached by \(\theta\mapsto \ii\pi-\theta\).  Hence the natural
analytic domain for a two-body amplitude is the physical strip
\[
  \mathcal S_{\rm phys}
  =
  \{\theta\in\C\mid 0<\operatorname{Im}\theta<\pi\}.
\]
For unequal masses \(m_a,m_b\),
\[
  s_{ab}(\theta)=m_a^2+m_b^2+2m_am_b\cosh\theta,
\]
and the same strip organizes the direct and crossed channels after the
species labels are crossed.

\begin{definition}[Factorized scattering bootstrap data]
\label{def:factorized-scattering-bootstrap-data}
Factorized scattering bootstrap data are an elastic factorized scattering
system of Definition~\ref{def:elastic-factorized-scattering-system} together
with the following extra data and hypotheses:
\begin{enumerate}[label=(\roman*)]
\item meromorphic continuation of each component
  \(S_{ab}^{cd}(\theta)\) to the physical strip
  \(\mathcal S_{\rm phys}\);
\item a specified finite or locally finite list of simple-pole locations
  \(\theta=\ii u_{ab}^{e}\), \(0<u_{ab}^{e}<\pi\), labelled by proposed
  direct-channel bound-state species \(e\);
\item finite-dimensional multiplicity spaces for these bound-state channels
  and intertwiners \(\Gamma_{ab}^{e,\alpha}\) specifying residues;
\item horizontal-strip growth bounds stated as follows: for every closed
  substrip \(K\subset\mathcal S_{\rm phys}\) whose distance from the declared
  pole set is positive, there are constants \(C_K,N_K\) such that
  \[
    \|S_{ab}^{cd}(\theta)\|
    \le C_K(1+|\operatorname{Re}\theta|)^{N_K},
    \qquad \theta\in K.
  \]
  Near each declared simple pole the corresponding principal part is
  subtracted before applying this estimate.  When a thermodynamic
  Bethe-ansatz kernel
  \(\phi_{ab}(\theta)=-\ii\partial_\theta\log S_{ab}(\theta)\) is used later,
  the datum must also specify the branch and the additional integrability
  hypothesis on \(\phi_{ab}\), usually \(L^1(\mathbb R)\) after removal of
  explicitly treated CDD or bound-state factors.
\end{enumerate}
\end{definition}

The growth condition in part (iv) is part of the datum, not a placeholder for
later arguments.  Stronger estimates may be imposed in form-factor or TBA
applications, but any such strengthening is an additional hypothesis on top
of the polynomial substrip bound displayed above.

The mass of a direct-channel bound state at \(\theta=\ii u_{ab}^{e}\) is
fixed by the pole location:
\[
  m_e^2
  =
  m_a^2+m_b^2+2m_am_b\cos u_{ab}^{e}.
\]
Indeed, write the incoming momenta as
\[
  p_a(\theta_a)=m_a(\cosh\theta_a,\sinh\theta_a),
  \qquad
  p_b(\theta_b)=m_b(\cosh\theta_b,\sinh\theta_b),
\]
with \(\theta=\theta_a-\theta_b\).  Then
\[
  -\bigl(p_a+p_b\bigr)^2
  =
  m_a^2+m_b^2+2m_am_b\cosh(\theta_a-\theta_b)
  =
  m_a^2+m_b^2+2m_am_b\cosh\theta .
\]
Analytic continuation to the pole \(\theta=\ii u_{ab}^{e}\) gives
\(\cosh(\ii u_{ab}^{e})=\cos u_{ab}^{e}\), hence the displayed formula.  A
one-particle vector of mass \(m_e\) in a Hilbert-space QFT is additional data
to be constructed or assumed.  The pole and residue record the on-shell
criterion that such a vector would have to satisfy; a physical-strip pole
with the wrong residue sign, or a pole related by crossing rather than the
direct-channel residue, is not by itself a new stable direct-channel
particle.

\section{Residues and Fusion}

Let \(V_a\) denote the internal one-particle space of species \(a\).  A
simple pole at \(\theta=\ii u_{ab}^{e}\) is interpreted as a direct-channel
bound-state pole only after one specifies a finite-dimensional multiplicity
space \(M_{ab}^{e}\) and maps
\[
  \Gamma_{e,\alpha}^{ab}\colon V_e\longrightarrow V_a\otimes V_b,
  \qquad
  \Gamma_{ab}^{e,\alpha}\colon V_a\otimes V_b\longrightarrow V_e,
  \qquad \alpha\in M_{ab}^{e},
\]
normalized by
\[
  \Gamma_{ab}^{e,\alpha}\Gamma_{e,\beta}^{ab}
  =
  \delta^\alpha{}_\beta\,\mathbf 1_{V_e}.
\]
With this convention the residue condition is the operator identity
\begin{equation}
  -\ii\,\operatorname*{Res}_{\theta=\ii u_{ab}^{e}}S_{ab}(\theta)
  =
  \sum_{\alpha\in M_{ab}^{e}}
  \Gamma_{e,\alpha}^{ab}\Gamma_{ab}^{e,\alpha}
  \quad
  \text{on }V_a\otimes V_b .
  \label{eq:bound-state-residue-projection}
\end{equation}
In components this is the factorized residue formula with one map carrying
the incoming \(ab\)-indices to the \(e\)-channel and the other carrying the
\(e\)-channel back to outgoing \(ab\)-indices.  Positivity of the
Hilbert-space norm fixes the sign in
\eqref{eq:bound-state-residue-projection} after the one-particle
normalization has been chosen.

The fusing angles are not independent symbols.  They are defined by the
complex on-shell momentum identity
\[
  p_a(\theta+\ii\bar u_{ae}^{b})
  +
  p_b(\theta-\ii\bar u_{be}^{a})
  =
  p_e(\theta),
  \qquad
  \bar u_{ae}^{b}+\bar u_{be}^{a}=u_{ab}^{e}.
\]
Equivalently,
\[
  m_a\sin \bar u_{ae}^{b}=m_b\sin \bar u_{be}^{a},
  \qquad
  m_e=m_a\cos \bar u_{ae}^{b}+m_b\cos \bar u_{be}^{a}.
\]
These equations are obtained by expanding the complex boosts:
\[
\begin{aligned}
  p_a(\theta+\ii\alpha)+p_b(\theta-\ii\beta)
  &=
  \bigl(m_a\cos\alpha+m_b\cos\beta\bigr)
  (\cosh\theta,\sinh\theta) \\
  &\quad
  +\ii\bigl(m_a\sin\alpha-m_b\sin\beta\bigr)
  (\sinh\theta,\cosh\theta),
\end{aligned}
\]
with \(\alpha=\bar u_{ae}^{b}\) and \(\beta=\bar u_{be}^{a}\).  Equality to the
real on-shell vector \(p_e(\theta)\) forces the imaginary coefficient to vanish
and the real coefficient to be \(m_e\).  Squaring this real coefficient then
recovers the pole-location mass formula because
\[
\begin{aligned}
  m_e^2
  &=
  \bigl(m_a\cos\alpha+m_b\cos\beta\bigr)^2 \\
  &=
  m_a^2+m_b^2+2m_am_b\cos(\alpha+\beta),
\end{aligned}
\]
where the last step uses
\(m_a\sin\alpha=m_b\sin\beta\) to replace
\(m_a^2\sin^2\alpha+m_b^2\sin^2\beta\) by
\(2m_am_b\sin\alpha\sin\beta\).  Thus the fusing-angle data are not extra
kinematics beyond the pole location and the on-shell momentum identity; they
are the component form of that identity.

Fusion is the compatibility condition between the residue projection
\eqref{eq:bound-state-residue-projection} and scattering with an additional
particle \(k\).  With the written composition convention that the rightmost
map acts first, the bootstrap equation is the meromorphic identity
\[
  S_{ek}(\eta)
  =
  \sum_{\alpha\in M_{ab}^{e}}
  \Gamma_{ab}^{e,\alpha}\,
  S_{ak}(\eta+\ii\bar u_{ae}^{b})\,
  S_{bk}(\eta-\ii\bar u_{be}^{a})\,
  \Gamma_{e,\alpha}^{ab},
  \qquad
  \eta=\theta-\theta_k .
\]
The Yang--Baxter equation is the statement that fusing before or after an
allowed adjacent reordering of three particles gives the same meromorphic
operator.

\begin{proposition}[Fusion identity as a residue identity]
\label{prop:fusion-identity-residue-derivation}
Assume that the three-particle scattering amplitude has meromorphic
factorization on the rapidity chamber under consideration and that the pair
\((a,b)\) has a simple direct-channel pole at
\(\theta_a-\theta_b=\ii u_{ab}^e\) with residue
\eqref{eq:bound-state-residue-projection}.  Assume also that, on the small
contour used below, no pole of the \(ak\) or \(bk\) two-body factors is met.
Choose the bound-state rapidity \(\theta\) so that the two constituent
rapidities are
\[
  \theta_a=\theta+\ii\bar u_{ae}^{b},
  \qquad
  \theta_b=\theta-\ii\bar u_{be}^{a},
  \qquad
  \bar u_{ae}^{b}+\bar u_{be}^{a}=u_{ab}^{e}.
\]
Then the residue of the factorized three-particle amplitude at the
\((a,b)\)-pole gives
\[
  S_{ek}(\theta-\theta_k)
  =
  \sum_{\alpha\in M_{ab}^{e}}
  \Gamma_{ab}^{e,\alpha}\,
  S_{ak}(\theta-\theta_k+\ii\bar u_{ae}^{b})\,
  S_{bk}(\theta-\theta_k-\ii\bar u_{be}^{a})\,
  \Gamma_{e,\alpha}^{ab}
\]
as an identity on the bound-state sector, after the one-particle
normalization used in the residue has been fixed.  If the multiplicity space
has dimension greater than one, the displayed expression is the trace over
the chosen residue basis; equivalently one may keep the corresponding
matrix-valued \(S_{ek}^{\alpha\beta}\).
\end{proposition}

\begin{proof}
Put
\[
  \eta=\theta-\theta_k,\qquad
  \zeta=\theta_a-\theta_b,
\]
and, near the pole \(\zeta=\ii u_{ab}^{e}\), parametrize the two constituent
rapidities by
\[
  \theta_a(\zeta)
  =
  \theta+\ii\bar u_{ae}^{b}
  +\frac{\zeta-\ii u_{ab}^{e}}{2},
  \qquad
  \theta_b(\zeta)
  =
  \theta-\ii\bar u_{be}^{a}
  -\frac{\zeta-\ii u_{ab}^{e}}{2}.
\]
Thus \(\theta_a(\zeta)-\theta_b(\zeta)=\zeta\), while at the pole the two
continued momenta add to \(p_e(\theta)\).  Choose a small positively oriented
circle \(C_e\) around \(\ii u_{ab}^{e}\) on which no singularity of the
factors \(S_{ak}(\theta_a(\zeta)-\theta_k)\) or
\(S_{bk}(\theta_b(\zeta)-\theta_k)\) is encountered.  The pole extraction is
the contour operation
\[
  -\ii\,\operatorname*{Res}_{\zeta=\ii u_{ab}^{e}}A(\zeta)
  =
  -\frac{\ii}{2\pi\ii}\oint_{C_e}A(\zeta)\,d\zeta .
\]

In the unresolved three-particle channel, factorization gives a meromorphic
amplitude of the form
\[
  A_{ab|k}(\zeta)
  =
  S_{ak}(\theta_a(\zeta)-\theta_k)\,
  S_{bk}(\theta_b(\zeta)-\theta_k)\,
  S_{ab}(\zeta),
\]
up to the common convention of writing operators in the order in which they
act.  The first two factors are holomorphic inside \(C_e\).  Hence Cauchy's
residue theorem and \eqref{eq:bound-state-residue-projection} give
\[
  -\ii\,\operatorname*{Res}_{\zeta=\ii u_{ab}^{e}}A_{ab|k}(\zeta)
  =
  S_{ak}(\eta+\ii\bar u_{ae}^{b})\,
  S_{bk}(\eta-\ii\bar u_{be}^{a})\,
  \sum_\alpha
  \Gamma_{e,\alpha}^{ab}\Gamma_{ab}^{e,\alpha}.
\]
Project this residue onto a fixed bound-state multiplicity component by
applying \(\Gamma_{ab}^{e,\alpha}\) on the outgoing \(ab\)-pair and
\(\Gamma_{e,\alpha}^{ab}\) on the incoming \(e\)-channel.  Using
\(\Gamma_{ab}^{e,\alpha}\Gamma_{e,\beta}^{ab}
=\delta^\alpha{}_\beta\mathbf 1_{V_e}\), the unresolved residue becomes
\[
  \Gamma_{ab}^{e,\alpha}\,
  S_{ak}(\eta+\ii\bar u_{ae}^{b})\,
  S_{bk}(\eta-\ii\bar u_{be}^{a})\,
  \Gamma_{e,\alpha}^{ab}.
\]
The same contour, evaluated after resolving the pole as a one-particle
intermediate state, is by definition the scattering amplitude
\(S_{ek}^{(\alpha)}(\eta)\) of that bound-state component with \(k\).  Since
both calculations are residues of the same meromorphic three-particle
amplitude and no other singularity crossed the contour, the two expressions
are equal.  Summing over the multiplicity basis gives the displayed fusion
identity.  The proof uses only meromorphic factorization, the residue
normalization, and the existence of the one-particle \(e\)-channel; it does
not use an off-shell Lagrangian.
\end{proof}

The noncommuting algebra in this residue step is finite-dimensional.  If
\(P_e=\sum_\alpha\Gamma_{e,\alpha}^{ab}\Gamma_{ab}^{e,\alpha}\) and
\[
  S_{ab}(\zeta)=\frac{\ii P_e}{\zeta-\ii u_{ab}^{e}}+S_{ab}^{\rm reg}(\zeta),
  \qquad
  B(\zeta)=
  S_{ak}(\theta_a(\zeta)-\theta_k)
  S_{bk}(\theta_b(\zeta)-\theta_k),
\]
with \(B\) holomorphic on \(C_e\), then
\[
  -\ii\operatorname*{Res}_{\zeta=\ii u_{ab}^{e}}B(\zeta)S_{ab}(\zeta)
  =
  B(\ii u_{ab}^{e})P_e .
\]
Projecting to the resolved bound-state coordinate gives
\[
  \Gamma_{ab}^{e,\alpha}B(\ii u_{ab}^{e})P_e\Gamma_{e,\beta}^{ab}
  =
  \Gamma_{ab}^{e,\alpha}B(\ii u_{ab}^{e})\Gamma_{e,\beta}^{ab}.
\]
Thus derivatives or regular coefficients of the \(ak\) and \(bk\) factors do
not enter this simple-pole residue, while the sign in
\eqref{eq:bound-state-residue-projection} fixes the sign of the fused
amplitude.  The companion script
\path{calculation-checks/integrable_scattering_bootstrap_checks.py} checks
this point in an exact rational finite-rank model with a nonorthogonal
projection and noncommuting holomorphic scattering factor.

\section{Minimal Solutions and CDD Factors}

For diagonal scattering, where \(S_{ab}^{cd}(\theta)=
\delta_a^c\delta_b^d S_{ab}(\theta)\), unitarity and crossing become
\[
  S_{ab}(\theta)S_{ab}(-\theta)=1,
  \qquad
  S_{ab}(\theta)=S_{\bar a b}(\ii\pi-\theta).
\]
A scalar meromorphic factor \(C(\theta)\) satisfying the same two equations
and having no assigned particle pole is a CDD factor.  Thus the analytic
constraints and the bound-state list determine the amplitude only after an
additional growth or minimality condition is specified.

\begin{definition}[Elementary scalar block]
\label{def:integrable-elementary-scalar-block}
For \(x\in\C\), the elementary scalar block is
\[
  [x]_\theta
  =
  \frac{\sinh\!\left((\theta+\ii\pi x)/2\right)}
       {\sinh\!\left((\theta-\ii\pi x)/2\right)} ,
\]
as a meromorphic function of \(\theta\).
\end{definition}

This block is the basic unitary zero-pole pair used in diagonal bootstrap
formulae.  Products of such blocks place zeros and poles at prescribed
locations.  The bootstrap construction problem is to choose the product and
possible CDD factors so that unitarity, crossing, physical-strip pole
assignments, residue signs, and asymptotic bounds hold simultaneously.

\begin{remark}[Elementary block algebra]
\label{rem:elementary-scalar-block-algebra}
The elementary scalar block satisfies
\[
  [x]_{\theta}[x]_{-\theta}=1.
\]
Its poles are
\[
  \theta=\ii\pi x+2\pi\ii n,\qquad n\in\mathbb Z,
\]
and its zeros are
\[
  \theta=-\ii\pi x+2\pi\ii n,\qquad n\in\mathbb Z .
\]
For real \(0<x<1\), the pole at \(\theta=\ii\pi x\) lies in the physical
strip.
\end{remark}

The displayed properties follow immediately from the zero set of
\(\sinh z\) and from \(\sinh(-z)=-\sinh z\).  The point worth keeping in the
main text is the residue sign.  At the physical-strip pole
\(\theta=\ii\pi x\), \(0<x<1\),
\[
  \operatorname*{Res}_{\theta=\ii\pi x}[x]_\theta
  =
  2\ii\sin(\pi x),
  \qquad
  -\ii\operatorname*{Res}_{\theta=\ii\pi x}[x]_\theta
  =
  2\sin(\pi x)>0 .
\]
Thus a single block has the residue sign compatible with a direct-channel
scalar bound-state pole in the convention
\eqref{eq:bound-state-residue-projection}.  Crossing is a separate
constraint.  The elementary relation
\[
  [x]_{\ii\pi-\theta}=-[1-x]_\theta
\]
shows that the crossing-symmetric scalar pair
\[
  \mathcal C_x(\theta):=[x]_\theta[1-x]_\theta
\]
is unitary and crossing invariant.  For \(0<x<1\) with \(x\ne1/2\), it has
physical-strip poles at \(\theta=\ii\pi x\) and
\(\theta=\ii\pi(1-x)\), but their residue signs are opposite:
\begin{equation}
  -\ii\operatorname*{Res}_{\theta=\ii\pi x}\mathcal C_x(\theta)
  =
  -2\tan(\pi x),
  \qquad
  -\ii\operatorname*{Res}_{\theta=\ii\pi(1-x)}\mathcal C_x(\theta)
  =
  2\tan(\pi x).
  \label{eq:crossing-pair-opposite-residue-signs}
\end{equation}
Equation~\eqref{eq:crossing-pair-opposite-residue-signs} is a useful
diagnostic: a CDD zero-pole pair can satisfy scalar unitarity and crossing
while still failing the positivity sign for one of its physical-strip poles.
The particle interpretation of a pole therefore uses the full residue datum,
not only its position in the strip.  The companion check
\path{calculation-checks/integrable_scattering_bootstrap_checks.py} verifies
the scalar-block identities and the two residue signs without printing the
elementary hyperbolic cancellation in the chapter.

\section{Hilbert-Space Boundary}

The bootstrap data are on-shell information.  A local QFT realization also
requires a Hilbert space, Wightman distributions, local operator domains, and
cluster properties.  In constructive treatments of factorizing models, wedge-
localized fields and modular localization provide one route from an
admissible \(S\)-matrix to local algebras.  The theorem boundary is precise:
analyticity and pole data enter as hypotheses for constructing wedge
operators and then proving that double-cone intersections are nontrivial.

\begin{openproblem}[Bootstrap completeness for local algebras]
\label{op:bootstrap-completeness-local-algebras}
Characterize the bootstrap data for which the resulting wedge-local
construction has nontrivial local algebras, a unique vacuum, asymptotic
completeness, and local fields whose form factors solve the equations of
Chapter~\ref{ch:integrable-factorized-scattering}.  This problem is the
operator-algebraic completion of the on-shell bootstrap.
\end{openproblem}
