\chapter{Form-Factor Bootstrap and Local Operators}
\label{ch:integrable-form-factor-bootstrap-local-operators}

\ConventionsLorentzian

The scattering specification of Chapters~\ref{ch:integrable-scattering-analyticity-bootstrap}
and~\ref{ch:integrable-yang-baxter-internal-symmetry} is on-shell.  Local QFT
contains more structure: local operator domains, Wightman distributions, and
matrix elements of local fields between finite-particle states.  In a
massive two-dimensional theory with factorized scattering, the first
operator-level object is the family of form factors.

\begin{definition}[Local-operator form-factor structure]
\label{def:form-factor-datum-local-operator}
Let \(\mathfrak R\) be a regular factorized scattering datum in the sense of
Definition~\ref{def:regular-two-body-r-system}.  A local-operator
form-factor structure over \(\mathfrak R\) is a tuple
\[
  \mathfrak F_{\mathcal O}
  =
  \bigl(
    \widehat{\mathcal O},
    \{F^{\mathcal O}_{a_1\ldots a_n}\}_{n\geq0,\ a_i\in\mathcal I},
    \{\omega_{\mathcal O,a}\}_{a\in\mathcal I},
    \mathcal P_{\rm kin},
    \mathcal P_{\rm bs},
    \mathcal G_{\mathcal O}
  \bigr)
\]
with the following entries.
\begin{enumerate}[label=\textnormal{(\roman*)}]
\item \(\widehat{\mathcal O}(f)\) is a scalar local or semi-local
  operator-valued distribution on a common dense scattering domain
  \(\mathcal D_{\rm scat}\subset\mathcal H\), with vacuum
  \(\Omega\in\mathcal D_{\rm scat}\).  Semi-locality means that the exchange
  of \(\widehat{\mathcal O}\) with a charged particle line may carry the
  monodromy specified in item (iii); ordinary locality is the special case
  with trivial monodromy for every species.
\item For every ordered species word
  \(\mathbf a=(a_1,\ldots,a_n)\), \(F^{\mathcal O}_{\mathbf a}\) is a
  meromorphic covector-valued function on the rapidity variables,
  \[
    F^{\mathcal O}_{\mathbf a}(\theta_1,\ldots,\theta_n)
    \in
    \bigl(V_{a_1}\otimes\cdots\otimes V_{a_n}\bigr)^\vee ,
  \]
  whose physical real boundary value in the chamber
  \(\theta_1>\cdots>\theta_n\) equals the vacuum-to-in-state matrix element
  of \(\widehat{\mathcal O}(0)\).  The word \(\mathbf a\) is part of the
  argument; when adjacent rapidities are analytically continued around one
  another, the corresponding species spaces are reordered by the scattering
  maps of \(\mathfrak R\).
\item \(\omega_{\mathcal O,a}\in\C^\times\) is the semi-locality monodromy
  acquired when a particle of species \(a\) is analytically continued once
  around the insertion of \(\mathcal O\).  For a strictly local bosonic scalar
  operator all \(\omega_{\mathcal O,a}\) are \(1\); for order/disorder fields
  in two-dimensional examples the same symbol records the controlled failure
  of single-valuedness.
\item \(\mathcal P_{\rm kin}\) consists of the annihilation-pole residue
  tensors determined by the antiparticle and crossing data of
  \(\mathfrak R\).  These tensors specify the residues at
  \(\theta_i-\theta_j=\ii\pi\) after the chosen tensor-ordering convention is
  fixed.
\item \(\mathcal P_{\rm bs}\) consists of the bound-state residue tensors
  inherited from the physical-strip poles of the two-body scattering maps.
  If no stable bound-state channel is present, this entry is empty.
\item \(\mathcal G_{\mathcal O}\) is a stated family of growth and
  boundary-value bounds.  It must be strong enough for each spectral series
  later claimed in the chapter to define the announced distribution or
  separated Euclidean function.  A form-factor structure without such bounds is
  only algebraic bootstrap input, not a reconstruction of a local field.
\end{enumerate}
The exchange, cyclicity, kinematic-pole, bound-state-pole, covariance, and
growth equations are developed below as constraints on
\(\mathfrak F_{\mathcal O}\).  The definition records the object to which
those constraints apply; it does not assert that a solution of the algebraic
equations has already constructed a Wightman QFT.
\end{definition}

\section{Finite-Particle Matrix Elements}

Let \(\mathfrak F_{\mathcal O}\) be a local-operator form-factor structure over
\(\mathfrak R\).  Thus \(\mathcal I\) is the set of stable particle species,
\(V_a\) their internal species spaces, and
\(S_{ab}(\theta):V_a\otimes V_b\to V_b\otimes V_a\) the two-body scattering
maps.  Ordered incoming states are denoted
\[
  \ket{\theta_1,a_1;\ldots;\theta_n,a_n}_{\rm in},
  \qquad
  \theta_1>\cdots>\theta_n .
\]
For a local scalar operator \(\widehat{\mathcal O}(x)\), the \(n\)-particle
form factor is the distributional boundary value
\[
  F^{\mathcal O}_{a_1\ldots a_n}(\theta_1,\ldots,\theta_n)
  =
  \bra{\Omega}\widehat{\mathcal O}(0)
  \ket{\theta_1,a_1;\ldots;\theta_n,a_n}_{\rm in}.
\]
If the species spaces are not one-dimensional, this is a covector on
\(V_{a_1}\otimes\cdots\otimes V_{a_n}\).  Translation covariance gives
\[
  \bra{\Omega}\widehat{\mathcal O}(x)
  \ket{\theta_1,a_1;\ldots;\theta_n,a_n}_{\rm in}
  =
  \exp\!\left[-\ii x_\mu\sum_{j=1}^n p_{a_j}^\mu(\theta_j)\right]
  F^{\mathcal O}_{a_1\ldots a_n}(\theta_1,\ldots,\theta_n).
\]
Thus the nontrivial information is at \(x=0\), while the distributional
character of \(\widehat{\mathcal O}\) enters through growth and boundary-value
conditions on \(F^{\mathcal O}\).

\section{Exchange Equation}

Analytic continuation in rapidities extends the ordered form factor to
domains in \(\C^n\).  Fix an adjacent pair and write
\[
  \mathcal C_+=\{\theta_i>\theta_{i+1}\},
  \qquad
  \mathcal C_-=\{\theta_{i+1}>\theta_i\}.
\]
The form factor has two distributional boundary values at the wall
\(\theta_i=\theta_{i+1}\), obtained by approaching the wall from the two
ordered chambers, or equivalently by the two standard \(i0\)-prescriptions
for the rapidity difference
\[
  F_+=\lim_{\epsilon\downarrow0}
  F(\ldots,\theta_i+\ii\epsilon,\theta_{i+1}-\ii\epsilon,\ldots),
  \qquad
  F_-=\lim_{\epsilon\downarrow0}
  F(\ldots,\theta_i-\ii\epsilon,\theta_{i+1}+\ii\epsilon,\ldots).
\]
The ordered scattering-state convention says that the corresponding
finite-particle boundary vectors are related by the adjacent two-body
scattering operator,
\begin{equation}
  \ket{\ldots;\theta_i,a_i;\theta_{i+1},a_{i+1};\ldots}_{+}
  =
  S_{a_i a_{i+1}}^{bc}(\theta_i-\theta_{i+1})
  \ket{\ldots;\theta_{i+1},c;\theta_i,b;\ldots}_{-}.
  \label{eq:ordered-state-boundary-exchange}
\end{equation}
This relation is the rapidity-space form of the scattering theorem for
ordered Haag--Ruelle states in a factorizing theory; locality is used
upstream in constructing the asymptotic fields and their scattering
relations, not as a one-line replacement for the boundary-value statement.
Pairing \eqref{eq:ordered-state-boundary-exchange} with the local-operator
covector gives Watson's exchange equation
\[
  F^{\mathcal O}_{\ldots a_i a_{i+1}\ldots}
  (\ldots,\theta_i,\theta_{i+1},\ldots)
  =
  F^{\mathcal O}_{\ldots a_{i+1} a_i\ldots}
  (\ldots,\theta_{i+1},\theta_i,\ldots)
  S_{a_i a_{i+1}}(\theta_i-\theta_{i+1}),
\]
where the \(S\)-matrix acts on the two corresponding species factors before
the covector is evaluated.

\begin{proposition}[Exchange from factorized scattering]
\label{prop:form-factor-exchange}
Assume that the finite-particle scattering states satisfy asymptotic
completeness in the massive sector and that the ordered in-state convention is
related across adjacent rapidity chambers by the factorized \(S\)-matrix.
Then every local operator whose finite-particle matrix elements admit
rapidity boundary values satisfies the exchange equation above.
\end{proposition}

\begin{proof}
Let \(g_+\) and \(g_-\) be compactly supported test functions in the two
neighboring rapidity chambers whose supports approach the same compact set on
the wall but do not meet any bound-state or threshold singularity.  Smearing
\eqref{eq:ordered-state-boundary-exchange} with these test functions gives an
identity of Hilbert-space vectors in the finite-particle scattering domain.
The identity is first proved for separated velocity supports by the
Haag--Ruelle construction of ordered in-states; analytic continuation in the
rapidity difference then supplies the displayed boundary values.  The
factorizing hypothesis identifies the change of ordered basis with the
adjacent two-body \(S\)-matrix and leaves all other tensor factors unchanged.

Applying the continuous distributional covector
\(\bra{\Omega}\widehat{\mathcal O}(0)\) to both sides gives the equality of
smeared form-factor boundary values.  Since this holds for all such test
functions, the equality holds as a distributional boundary-value identity in
the rapidity variables.  No off-shell Lagrangian is used.  The hypotheses are
the existence of the finite-particle scattering states, their
boundary-value analyticity in rapidity chambers, and the factorized
two-body change of ordered basis.
\end{proof}

\section{Cyclicity, Locality, and Semi-Locality}

The cyclicity equation is a crossing-locality statement.  It is not obtained
from the two-body \(S\)-matrix alone.  The analytic input is the same as in
the Wightman reconstruction chapters: the spectrum condition gives
forward-tube analyticity of vacuum matrix elements, and locality identifies
boundary values after permuting spacelike separated insertions.  In rapidity
variables the complex Lorentz transformation
\[
  \theta\longmapsto\theta+\ii\pi
\]
changes \(p_a(\theta)\) to \(-p_a(\theta)\), and hence crosses an incoming
particle to the outgoing antiparticle channel.  A second shift by \(\ii\pi\)
crosses it back to the incoming side after it has passed through the local
operator.

\begin{proposition}[Cyclicity from crossing and locality]
\label{prop:form-factor-cyclicity-crossing-locality}
Assume that the finite-particle matrix elements of a scalar local operator
\(\widehat{\mathcal O}\) are boundary values of Wightman functions satisfying
the spectrum condition, locality, and the one-particle crossing relation
obtained by analytic continuation through the physical rapidity strip.  Assume
that the path
\(\theta_1\mapsto\theta_1+2\pi\ii\) crosses no singularity other than the
standard crossing boundary and no bound-state pole.  Then
\[
  F^{\mathcal O}_{a_1a_2\ldots a_n}
  (\theta_1+2\pi\ii,\theta_2,\ldots,\theta_n)
  =
  F^{\mathcal O}_{a_2\ldots a_n a_1}
  (\theta_2,\ldots,\theta_n,\theta_1).
\]
If \(\widehat{\mathcal O}\) is semi-local with respect to species \(a_1\),
the right side is multiplied by the declared monodromy endomorphism
\(\omega_{\mathcal O,a_1}\).
\end{proposition}

\begin{proof}
For separated velocity supports, represent the incoming particle of species
\(a_1\) by a Haag--Ruelle wave packet and write the matrix element as a
boundary value of a Wightman distribution with one interpolating field for
that particle.  The joint spectrum condition implies that this distribution
extends holomorphically when the imaginary parts of successive spacetime
differences lie in the future cone.  Fourier transforming the first wave
packet and using \(p_a(\theta+\ii\pi)=-p_a(\theta)\), continuation
\(\theta_1\mapsto\theta_1+\ii\pi\) converts the incoming \(a_1\)-particle
into an outgoing \(\bar a_1\)-particle.  In bracket notation this is the
crossing boundary value
\[
  F^{\mathcal O}_{a_1a_2\ldots a_n}
  (\theta_1+\ii\pi,\theta_2,\ldots,\theta_n)
  =
  {}_{\rm out}\!\bra{\theta_1,\bar a_1}
  \widehat{\mathcal O}(0)
  \ket{\theta_2,a_2;\ldots;\theta_n,a_n}_{\rm in},
\]
with the charge-conjugation map suppressed in the notation.

Continue once more by \(\ii\pi\).  At the boundary of the tube one may choose
the spacetime supports of the crossed particle and of
\(\widehat{\mathcal O}\) to be spacelike separated before the wave-packet
limit is taken.  Local commutativity then permits the crossed particle
interpolating field to be moved past \(\widehat{\mathcal O}\).  For a scalar
local operator this gives no extra factor; for a semi-local operator it gives
the declared monodromy \(\omega_{\mathcal O,a_1}\).  After the second crossing
the particle is again incoming, and the ordered-state convention places its
rapidity at the right end of the ordered list.  Thus the boundary value is
the displayed cyclic form factor.  The equality is distributional because it
was derived after smearing in rapidity and then unsmearing by uniqueness of
boundary values.  The excluded pole crossings are precisely the cases where
additional residue terms must be added.
\end{proof}

\section{Kinematic Pole Equation}

Let \(\bar a\) denote the antiparticle species of \(a\).  The pair
\((\bar a,a)\) at rapidity difference \(\ii\pi\) has total momentum zero and
can annihilate into the vacuum.  Consequently a local form factor may have a
kinematic pole at
\[
  \theta_{\bar a}=\theta_a+\ii\pi .
\]
For a scalar local operator the residue has the structural form
\begin{equation}
  -\ii\,\operatorname*{Res}_{\theta'=\theta+\ii\pi}
  F^{\mathcal O}_{\bar a\,a\,a_1\ldots a_n}
  (\theta',\theta,\theta_1,\ldots,\theta_n)
  =
  \left(
    1-\prod_{j=1}^n
    S_{a a_j}(\theta-\theta_j)
  \right)
  F^{\mathcal O}_{a_1\ldots a_n}(\theta_1,\ldots,\theta_n),
\label{eq:form-factor-kinematic-pole}
\end{equation}
with tensor contractions and charge-conjugation maps inserted according to
the species spaces.  The product is ordered by moving the particle \(a\)
through the \(n\) remaining particles before annihilation.

\begin{proposition}[Kinematic annihilation pole]
\label{prop:form-factor-kinematic-pole-derived}
Assume the hypotheses of
Proposition~\ref{prop:form-factor-cyclicity-crossing-locality}, asymptotic
completeness in the finite-particle sector under consideration, factorized
scattering for ordered states, and the standard rapidity normalization of
one-particle states.  Assume further that the only singularity encountered at
\(\theta'=\theta+\ii\pi\) is the vacuum annihilation of the adjacent pair
\((\bar a,a)\).  Then the singular part of the form factor is
\[
  F^{\mathcal O}_{\bar a\,a\,a_1\ldots a_n}
  (\theta',\theta,\theta_1,\ldots,\theta_n)
  =
  \frac{\ii}{\theta'-\theta-\ii\pi}
  \left(
    1-\prod_{j=1}^nS_{a a_j}(\theta-\theta_j)
  \right)
  F^{\mathcal O}_{a_1\ldots a_n}
  +\text{holomorphic},
\]
with the same tensor-ordering convention as in
\eqref{eq:form-factor-kinematic-pole}.  Equivalently,
\eqref{eq:form-factor-kinematic-pole} holds.
\end{proposition}

\begin{proof}
Set \(z=\theta'-\theta-\ii\pi\).  The coefficient of the crossed
one-particle contraction is fixed by matching its two boundary values to the
rapidity delta function, not by an independent sign convention.  In the
rapidity normalization, valid for any stable species \(b\),
\[
  \langle \theta',b|\theta,b\rangle=2\pi\delta(\theta'-\theta),
\]
the contraction term must have boundary-value jump \(2\pi\delta(x)\) when
\(x=\operatorname{Re}z\).  The distributional identity
\[
  \frac{1}{x+\ii0}-\frac{1}{x-\ii0}
  =
  -2\pi\ii\,\delta(x)
\]
therefore selects
\[
  \frac{\ii}{z}.
\]
Indeed,
\(\ii(x+\ii0)^{-1}-\ii(x-\ii0)^{-1}=2\pi\delta(x)\), exactly the
one-particle contraction coefficient.  This is the rapidity form of the
mass-shell phase-space pole: near \(z=0\), the continued total momentum
\(p_{\bar a}(\theta+\ii\pi+z)+p_a(\theta)\) vanishes linearly in \(z\), and the
chosen positive-energy boundary value is the one just normalized.  Thus direct
annihilation of the adjacent \((\bar a,a)\)-pair into the vacuum contributes
\[
  \frac{\ii}{z}
  F^{\mathcal O}_{a_1\ldots a_n}(\theta_1,\ldots,\theta_n).
  \tag{direct}
\]

There is a second boundary value of the same meromorphic matrix element.
Use Watson exchange to move the particle \(a\) through the spectator
particles \(a_1,\ldots,a_n\) before performing the annihilation.  This
ordered-basis change contributes the ordered product
\[
  \prod_{j=1}^n S_{a a_j}(\theta-\theta_j).
\]
The annihilation contour now approaches the same Cauchy pole from the
opposite side of the rapidity ordering.  Equivalently, after applying
cyclicity, the crossed particle encircles the local operator and returns to
the incoming side with the opposite orientation for the small circle around
\(z=0\).  In local coordinates this is the statement that the reverse
ordering uses
\[
  z_{\rm rev}
  =
  \theta-\theta'+\ii\pi
  =
  -z,
\]
so a positively oriented small circle in \(z_{\rm rev}\) is a negatively
oriented small circle in \(z\).  Therefore the residue of the same
annihilation pole enters with the opposite sign.  The second singular
contribution is
\[
  -\frac{\ii}{z}
  \left(\prod_{j=1}^n S_{a a_j}(\theta-\theta_j)\right)
  F^{\mathcal O}_{a_1\ldots a_n}(\theta_1,\ldots,\theta_n).
  \tag{scattered}
  \]
For a scalar local operator there is no extra semi-locality phase; a
semi-local operator would multiply this second contribution by its monodromy
factor.  Adding the two boundary contributions gives the displayed singular
part.  Taking the residue at \(z=0\) gives
\[
  -\ii\,\operatorname*{Res}_{z=0}
  \frac{\ii}{z}
  \left(1-\prod_{j=1}^nS_{a a_j}(\theta-\theta_j)\right)
  F_n^{\mathcal O}
  =
  \left(1-\prod_{j=1}^nS_{a a_j}(\theta-\theta_j)\right)
  F_n^{\mathcal O}.
\]
The proof is local in the rapidity variables; additional bound-state poles or
threshold singularities on the deformation path would add further residue
terms and are excluded by hypothesis.
\end{proof}

\section{Free Majorana Examples}

The form-factor axioms become concrete already in the massive Ising field
theory.  In the fermionic particle basis there is one self-conjugate particle
of mass \(m\), and the diagonal two-body scattering function is the constant
\[
  S(\theta)=-1 .
\]
The constant phase is physically important: the ordered rapidity variables
are still bosonic coordinates on asymptotic particle states, but exchanging
two particles multiplies the finite-particle basis by \(-1\).  Thus a scalar
form factor must be antisymmetric in rapidity variables when it has an even
number of Majorana particles.

\begin{example}[Energy-density form factor]
\label{ex:free-majorana-energy-form-factor}
Fix a scalar normalization constant \(\kappa_{\varepsilon}\).  The massive
Majorana energy-density operator \(\varepsilon\) may be normalized by
\[
  F_0^\varepsilon=0,\qquad
  F_2^\varepsilon(\theta_1,\theta_2)
  =
  \kappa_{\varepsilon}
  \sinh\!\left(\frac{\theta_1-\theta_2}{2}\right),
  \qquad
  F_n^\varepsilon=0\quad(n\neq2).
\]
The exchange equation follows because
\[
  \sinh\!\left(\frac{\theta_2-\theta_1}{2}\right)(-1)
  =
  \sinh\!\left(\frac{\theta_1-\theta_2}{2}\right).
\]
Cyclicity follows from
\[
  \sinh\!\left(\frac{\theta_1+2\pi\ii-\theta_2}{2}\right)
  =
  -\sinh\!\left(\frac{\theta_1-\theta_2}{2}\right)
  =
  \sinh\!\left(\frac{\theta_2-\theta_1}{2}\right).
\]
The kinematic-pole equation is also satisfied: the displayed form factors
are entire, and the right side vanishes because \(F_0^\varepsilon=0\) for no
spectators and because \(1-S^2=0\) for two spectators.  This example is
finite rather than recursive; it checks that the sign conventions of Watson
exchange and \(2\pi\ii\) cyclicity are compatible with a familiar local
quadratic field.
\end{example}

\begin{proposition}[Two-particle reconstruction for the Majorana energy operator]
\label{prop:majorana-energy-two-particle-reconstruction}
Use the one-particle normalization
\[
  \braket{\theta'}{\theta}=2\pi\delta(\theta'-\theta),
  \qquad
  p^\mu(\theta)=m(\cosh\theta,\sinh\theta),
  \qquad
  p(\theta)^2=-m^2 .
\]
Let \(W_\varepsilon(x)=\langle\Omega|\varepsilon(x)\varepsilon(0)|\Omega\rangle\)
at noncoincident points, and define the Fourier transform by
\[
  \widetilde W_\varepsilon(P)
  =
  \int_{\R^{1,1}}\dd^2x\,e^{\ii P\cdot x}W_\varepsilon(x),
  \qquad
  W_\varepsilon(x)
  =
  \int\frac{\dd^2P}{(2\pi)^2}\,
  e^{-\ii P\cdot x}\widetilde W_\varepsilon(P).
\]
With \(s=-P^2\), the two-particle form factor in
Example~\ref{ex:free-majorana-energy-form-factor} gives
\begin{equation}
  \widetilde W_\varepsilon(P)
  =
  \theta(P^0)\theta(s-4m^2)\,
  \frac{|\kappa_\varepsilon|^2}{2m^2}
  \sqrt{1-\frac{4m^2}{s}}
  \label{eq:majorana-energy-two-particle-spectral-density}
\end{equation}
in this Fourier convention.  Equivalently, for Euclidean time separation
\(r>0\),
\begin{equation}
  G_\varepsilon(r)
  =
  \langle\varepsilon(r)\varepsilon(0)\rangle_E
  =
  \frac{|\kappa_\varepsilon|^2}{2\pi^2}
  \int_0^\infty \dd\alpha\,
  \sinh^2\!\left(\frac{\alpha}{2}\right)
  K_0\!\left(2mr\cosh\!\left(\frac{\alpha}{2}\right)\right),
  \label{eq:majorana-energy-euclidean-bessel-reconstruction}
\end{equation}
where, for \(z>0\),
\[
  K_0(z)=\int_0^\infty e^{-z\cosh u}\,\dd u .
\]
\end{proposition}

\begin{proof}
The finite-particle resolution of the identity in the one-species sector is
\[
  \mathbf 1_{\rm fin}
  =
  \sum_{n\ge0}\frac{1}{n!}
  \int_{\R^n}\prod_{j=1}^n\frac{\dd\theta_j}{2\pi}\,
  \ket{\theta_1,\ldots,\theta_n}_{\rm in}
  {}_{\rm in}\!\bra{\theta_1,\ldots,\theta_n}.
\]
For the free Majorana energy operator only the two-particle term is present.
Therefore, at noncoincident Lorentzian separation,
\[
  W_\varepsilon(x)
  =
  \frac12
  \int\frac{\dd\theta_1}{2\pi}\frac{\dd\theta_2}{2\pi}\,
  \left|F_2^\varepsilon(\theta_1,\theta_2)\right|^2
  e^{-\ii x\cdot(p(\theta_1)+p(\theta_2))}.
\]
The factor \(1/2\) is not optional: the rapidity integration is over all
ordered labels, while the two Majorana particles are identical.  Fourier
transformation gives
\[
  \widetilde W_\varepsilon(P)
  =
  \frac12
  \int\frac{\dd\theta_1}{2\pi}\frac{\dd\theta_2}{2\pi}\,
  (2\pi)^2\delta^{(2)}(P-p(\theta_1)-p(\theta_2))
  \left|F_2^\varepsilon(\theta_1,\theta_2)\right|^2 .
\]
The support condition is already visible: \(P\) must be future directed and
\(s=-P^2\ge4m^2\).

Because the integrand is Lorentz invariant except for the explicit vector
\(P\), evaluate it in the frame \(P=(M,0)\), \(M=\sqrt s\).  Put
\[
  \Theta=\frac{\theta_1+\theta_2}{2},
  \qquad
  \alpha=\theta_1-\theta_2 .
\]
Then
\[
  p(\theta_1)+p(\theta_2)
  =
  2m\cosh\!\left(\frac{\alpha}{2}\right)
  (\cosh\Theta,\sinh\Theta),
\]
and the equations imposed by the delta function have the two roots
\[
  \Theta=0,\qquad
  \alpha=\pm\alpha_s,\qquad
  \cosh\!\left(\frac{\alpha_s}{2}\right)=\frac{\sqrt s}{2m}.
\]
The absolute value of the Jacobian of
\((\Theta,\alpha)\mapsto(p^0_1+p^0_2,p^1_1+p^1_2)\) at either root is
\[
  \left|
  \det
  \begin{pmatrix}
  \partial_\Theta P^0 & \partial_\alpha P^0\\
  \partial_\Theta P^1 & \partial_\alpha P^1
  \end{pmatrix}
  \right|
  =
  \sqrt s\,m\,\sinh\!\left(\frac{\alpha_s}{2}\right).
\]
The two roots in \(\alpha\) cancel the identical-particle factor \(1/2\).
Using
\[
  \left|F_2^\varepsilon\right|^2
  =
  |\kappa_\varepsilon|^2
  \sinh^2\!\left(\frac{\alpha_s}{2}\right),
  \qquad
  \sinh^2\!\left(\frac{\alpha_s}{2}\right)
  =
  \frac{s-4m^2}{4m^2},
\]
one obtains \eqref{eq:majorana-energy-two-particle-spectral-density}.
The threshold behavior is therefore \(\sqrt{s-4m^2}\), not a pole: the
two-body phase-space square-root singularity is softened by the antisymmetric
energy-density form factor, which vanishes at equal rapidities.

For the Euclidean formula, take a positive Euclidean time separation \(r\).
The positive-energy spectral term is the analytic continuation of the same
finite-particle contribution:
\[
  G_\varepsilon(r)
  =
  \frac12
  \int\frac{\dd\theta_1}{2\pi}\frac{\dd\theta_2}{2\pi}\,
  |\kappa_\varepsilon|^2
  \sinh^2\!\left(\frac{\theta_1-\theta_2}{2}\right)
  e^{-mr(\cosh\theta_1+\cosh\theta_2)} .
\]
Changing variables to \((\Theta,\alpha)\) has unit Jacobian and gives
\[
  G_\varepsilon(r)
  =
  \frac{|\kappa_\varepsilon|^2}{8\pi^2}
  \int_{\R}\dd\Theta\int_{\R}\dd\alpha\,
  \sinh^2\!\left(\frac{\alpha}{2}\right)
  e^{-2mr\cosh(\alpha/2)\cosh\Theta}.
\]
By the integral definition of \(K_0\),
\[
  \int_{-\infty}^{\infty}
  e^{-2mr\cosh(\alpha/2)\cosh\Theta}\,\dd\Theta
  =
  2K_0\!\left(2mr\cosh\!\left(\frac{\alpha}{2}\right)\right).
\]
The remaining integrand is even in \(\alpha\), giving
\eqref{eq:majorana-energy-euclidean-bessel-reconstruction}.
All equalities in this proof are separated-point equalities.  Possible
contact terms at \(x=0\) are not part of the finite-particle spectral
calculation.
\end{proof}

\begin{proposition}[Free-field reconstruction status for the Majorana energy density]
\label{prop:majorana-energy-local-reconstruction-certificate}
In the massive Ising theory realized as the local free Majorana field, let
\(\varepsilon(x)\) be the normal-ordered quadratic energy-density Wick
polynomial, with the normalization of
Example~\ref{ex:free-majorana-energy-form-factor}.  Then the finite-particle
domain of the CAR Fock space gives a theorem-level construction of this local
observable with the following on-shell shadow.
\[
  S(\theta)=-1,\qquad
  \langle\Omega|\varepsilon(0)
  |\theta_1,\ldots,\theta_n\rangle_{\rm in}
  =
  \begin{cases}
  \kappa_\varepsilon
  \sinh\!\left((\theta_1-\theta_2)/2\right),& n=2,\\
  0,& n\neq2 .
  \end{cases}
\]
Consequently the separated two-point function of \(\varepsilon\) is exactly
the two-particle spectral integral computed in
Proposition~\ref{prop:majorana-energy-two-particle-reconstruction}; there is
no uncomputed higher-particle tail for this observable.

The status statement is part of the assertion.  The local field, common
domain, positivity, spacelike locality of the even Wick polynomial, and
Fock completeness of the one-particle Majorana sector come from the
free-Majorana CAR construction.  The form-factor equations in
Example~\ref{ex:free-majorana-energy-form-factor} record the matrix elements
of that already constructed field; by themselves they would not constitute a
local-field reconstruction theorem.
\end{proposition}

\begin{proof}
The free massive Majorana field is built on antisymmetric Fock space from a
positive one-particle representation of the two-dimensional Poincare group.
The finite-particle vectors are a common invariant core for smeared Wick
polynomials.  The CAR anticommutator has causal support, so even Wick
polynomials, in particular the energy density, commute at spacelike
separation.  Positivity and Fock completeness of the one-particle Majorana
sector are inherited from the Fock construction.

In this representation the in/out creation operators differ only by the
fermionic exchange convention, hence the scalar factorized scattering
function is \(S=-1\).  Since \(\varepsilon\) is normal ordered and quadratic,
its vacuum-to-\(n\)-particle matrix element vanishes for \(n=0\), for odd
\(n\) by fermion parity, and for every \(n>2\) by Wick degree.  The only
remaining creation part is the two-particle bilinear.  Lorentz covariance,
antisymmetry under particle exchange, and the chosen scalar normalization fix
it to
\(\kappa_\varepsilon\sinh((\theta_1-\theta_2)/2)\).  Substituting this single
nonzero channel into the finite-particle resolution of the identity is exactly
the calculation of
Proposition~\ref{prop:majorana-energy-two-particle-reconstruction}.
\end{proof}

The spin and disorder fields of the Ising theory are more instructive because
their form factors form infinite families and because their locality relative
to the Majorana particle is a branch-field structure with an explicit monodromy.
Let
\[
  P_N(\theta_1,\ldots,\theta_N)
  =
  \prod_{1\le i<j\le N}
  \tanh\!\left(\frac{\theta_i-\theta_j}{2}\right),
  \qquad P_0=P_1=1 .
\]
The function \(P_N\) is antisymmetric and satisfies
\[
  P_N(\theta_1+2\pi\ii,\theta_2,\ldots,\theta_N)
  =
  P_N(\theta_1,\ldots,\theta_N),
  \qquad
  P_N(\theta_2,\ldots,\theta_N,\theta_1)
  =
  (-1)^{N-1}P_N(\theta_1,\ldots,\theta_N).
\]
Thus a nonzero product family with even \(N\) solves cyclicity with
semi-locality phase \(-1\), while a nonzero product family with odd \(N\)
solves cyclicity with phase \(+1\).  This phase is part of the operator
sector.  It is not determined by the diagonal scattering function alone.

In the thermally perturbed Ising field theory, the sign of the Majorana mass
selects which of the lattice order/disorder scaling fields has a nonzero
vacuum expectation value.  We write \(\sigma_+\) for the spin field in the
ordered convention in which
\[
  \bar\sigma=\langle\Omega|\sigma_+(0)|\Omega\rangle\ne0 .
\]
With the conformal normalization
\(\langle\sigma(x)\sigma(0)\rangle\sim |x|^{-1/4}\) at short distance, the
standard infinite-plane normalization is
\[
  \bar\sigma
  =
  2^{1/12}e^{-1/8}A_{\rm G}^{3/2}|m|^{1/8},
\]
where \(A_{\rm G}\) is the Glaisher constant.  The derivation below uses only
the symbol \(\bar\sigma\), because the form-factor equations determine the
rapidity dependence and relative phases, not the ultraviolet normalization of
the spin field.

\paragraph{Even spin-field form factors in the free massive fermion.}
Let the Majorana particle have \(S=-1\), and let \(\sigma_+\) be a scalar
branch field with semi-locality phase \(-1\) relative to the Majorana
particle.  The rapidity dependence selected by the form-factor equations is
the even family
\begin{equation}
  F_{2k}^{\sigma_+}(\theta_1,\ldots,\theta_{2k})
  =
  \bar\sigma\,\ii^k P_{2k}(\theta_1,\ldots,\theta_{2k}),
  \qquad
  F_{2k+1}^{\sigma_+}=0 ,
  \label{eq:ising-spin-even-form-factors}
\end{equation}
It satisfies Watson exchange, semi-local cyclicity
\[
  F_{2k}^{\sigma_+}(\theta_1+2\pi\ii,\theta_2,\ldots,\theta_{2k})
  =
  -F_{2k}^{\sigma_+}(\theta_2,\ldots,\theta_{2k},\theta_1),
\]
and the semi-local annihilation-pole equation
\[
  -\ii\,\operatorname*{Res}_{\theta'=\theta+\ii\pi}
  F_{n+2}^{\sigma_+}(\theta',\theta,\eta_1,\ldots,\eta_n)
  =
  \left(1+(-1)^n\right)
  F_n^{\sigma_+}(\eta_1,\ldots,\eta_n).
\]
In particular,
\[
  F_0^{\sigma_+}=\bar\sigma,\qquad
  F_2^{\sigma_+}(\theta_1,\theta_2)
  =
  \ii\bar\sigma
  \tanh\!\left(\frac{\theta_1-\theta_2}{2}\right).
\]
Away from coincident bra and ket rapidities, crossing \(K\) outgoing
particles gives the connected analytic part of the mixed matrix element
\begin{align}
  &{}_{\rm out}\!\bra{\beta_1,\ldots,\beta_K}
  \sigma_+(0)
  \ket{\beta'_1,\ldots,\beta'_N}_{\rm in}^{\rm conn}
  \notag\\
  &\qquad =
  \bar\sigma\,\ii^{(K+N)/2}
  \prod_{1\le i<j\le K}
  \tanh\!\left(\frac{\beta_i-\beta_j}{2}\right)
  \prod_{1\le p<q\le N}
  \tanh\!\left(\frac{\beta'_p-\beta'_q}{2}\right)
  \prod_{\substack{1\le s\le K\\1\le t\le N}}
  \coth\!\left(\frac{\beta_s-\beta'_t}{2}\right),
  \label{eq:ising-spin-mixed-fz-form-factor}
\end{align}
for \(K+N\) even; it vanishes for \(K+N\) odd.  The superscript
``conn'' reminds us that the full distributional matrix element also has
disconnected delta-function contractions when some bra rapidity coincides
with a ket rapidity.

The verification is the following explicit product calculation.  Watson
exchange follows because \(P_N\) is antisymmetric and \(S=-1\).
For \(N=2k\), the two displayed identities for \(P_N\) give
\[
  P_N(\theta_1+2\pi\ii,\theta_2,\ldots,\theta_N)
  =
  -P_N(\theta_2,\ldots,\theta_N,\theta_1),
\]
which is exactly cyclicity with semi-locality phase \(-1\).

For the pole equation, set \(\theta'=\theta+\ii\pi+\epsilon\).  The
annihilating pair contributes
\[
  \tanh\!\left(\frac{\theta'-\theta}{2}\right)
  =
  \coth\!\left(\frac{\epsilon}{2}\right)
  =
  \frac{2}{\epsilon}+O(\epsilon),
\]
while each spectator pair gives
\[
  \tanh\!\left(\frac{\theta'-\eta_j}{2}\right)
  \tanh\!\left(\frac{\theta-\eta_j}{2}\right)
  \longrightarrow 1 .
\]
Hence
\[
  \operatorname*{Res}_{\theta'=\theta+\ii\pi}
  P_{n+2}(\theta',\theta,\eta_1,\ldots,\eta_n)
  =
  2P_n(\eta_1,\ldots,\eta_n).
\]
When \(n\) is even, \(F_n^{\sigma_+}=\bar\sigma \ii^{n/2}P_n\) and
\(F_{n+2}^{\sigma_+}=\bar\sigma \ii^{n/2+1}P_{n+2}\), so
\[
  -\ii\,\operatorname*{Res}F_{n+2}^{\sigma_+}
  =
  2F_n^{\sigma_+}.
\]
This is the displayed equation because \(1+(-1)^n=2\).  When \(n\) is odd,
both \(F_n^{\sigma_+}\) and \(F_{n+2}^{\sigma_+}\) vanish.

The crossed formula is a direct consequence of the same even family.  Move
\(K\) outgoing particles to the incoming side by the boundary-value
continuation \(\beta_s\mapsto\beta_s+\ii\pi\).  Products among two crossed
rapidities and among two uncrossed rapidities remain \(\tanh\)-products,
because \(\tanh(z+\pi\ii)=\tanh z\).  A crossed-uncrossed pair gives
\[
  \tanh\!\left(\frac{\beta_s+\ii\pi-\beta'_t}{2}\right)
  =
  \coth\!\left(\frac{\beta_s-\beta'_t}{2}\right).
\]
Substitution in \(F_{K+N}^{\sigma_+}\) gives
\eqref{eq:ising-spin-mixed-fz-form-factor}.  The exclusion of coincident
rapidities is necessary because crossing a particle onto an identical
particle in the other state also produces the ordinary Fock-space contraction
terms; those delta terms are separate from the meromorphic connected form
factor derived here.

The case \(K=N=1\) gives the convention-sensitive spin-field matrix element
used in the Fonseca--Zamolodchikov treatment:
\[
  {}_{\rm out}\!\bra{\beta_1}\sigma_+(0)\ket{\beta_2}_{\rm in}^{\rm conn}
  =
  \ii\bar\sigma
  \coth\!\left(\frac{\beta_1-\beta_2}{2}\right).
\]
In the present chapter this formula is the crossed boundary value of the
two-particle spin form factor above.\footnote{A
useful comparison point is P.~Fonseca and A.~Zamolodchikov,
``Ising field theory in a magnetic field: analytic properties of the free
energy'', arXiv:hep-th/0112167, and ``Ward identities and integrable
differential equations in the Ising field theory'', arXiv:hep-th/0309228.
The formulas here are rederived in the rapidity and crossing conventions of
this chapter.}

The dual odd product family is obtained by the same algebra with the opposite
cyclicity phase.  In a normalization \(v\in\C\) it is
\begin{equation}
  F_{2k+1}^{\Sigma}(\theta_1,\ldots,\theta_{2k+1})
  =
  v\,\ii^k
  \prod_{1\le i<j\le 2k+1}
  \tanh\!\left(\frac{\theta_i-\theta_j}{2}\right),
  \qquad
  F_{2k}^{\Sigma}=0 .
  \label{eq:ising-order-form-factors}
\end{equation}
The constant \(v=F_1^\Sigma\) is the one-particle normalization in this
sector.  Which product family is called order or disorder after matching to
the lattice scaling limit depends on the sign convention for the Majorana
mass and on the Kramers--Wannier identification.  The mathematical point here
is independent of that naming convention: the parity of the nonzero
form-factor family is fixed by the declared cyclicity phase.

\paragraph{Odd-family bootstrap verification.}
For the odd product family \eqref{eq:ising-order-form-factors}, the three
bootstrap checks are elementary but convention-sensitive.  The family satisfies
Watson exchange, the local cyclicity equation for odd particle number, and the
kinematic-pole equation with \(S=-1\):
\[
  -\ii\,\operatorname*{Res}_{\theta'=\theta+\ii\pi}
  F_{n+2}^{\Sigma}(\theta',\theta,\eta_1,\ldots,\eta_n)
  =
  \left(1-(-1)^n\right)F_n^\Sigma(\eta_1,\ldots,\eta_n)
\]
for every odd \(n\).  For even \(n\), both sides vanish because
\(F_n^\Sigma=F_{n+2}^\Sigma=0\).

Interchanging adjacent variables changes exactly one factor by a sign and
permutes the rest, so \(P_N\) is antisymmetric.  Since \(S=-1\), this is
precisely Watson's equation.

For \(N\) odd,
\[
  P_N(\theta_1+2\pi\ii,\theta_2,\ldots,\theta_N)
  =
  P_N(\theta_1,\theta_2,\ldots,\theta_N),
\]
because \(\tanh(z+\pi\ii)=\tanh z\).  Moving \(\theta_1\) from the first
slot to the last reverses its order relative to \(N-1\) variables and hence
multiplies the antisymmetric product by \((-1)^{N-1}=1\).  This proves
cyclicity for the odd family.

It remains to compute the annihilation residue.  Put
\[
  \theta'=\theta+\ii\pi+\epsilon .
\]
The pair factor is
\[
  \tanh\!\left(\frac{\theta'-\theta}{2}\right)
  =
  \tanh\!\left(\frac{\ii\pi+\epsilon}{2}\right)
  =
  \coth\!\left(\frac{\epsilon}{2}\right)
  =
  \frac{2}{\epsilon}+O(\epsilon).
\]
For each spectator rapidity \(\eta_j\),
\[
  \tanh\!\left(\frac{\theta'-\eta_j}{2}\right)
  \tanh\!\left(\frac{\theta-\eta_j}{2}\right)
  \longrightarrow
  \coth\!\left(\frac{\theta-\eta_j}{2}\right)
  \tanh\!\left(\frac{\theta-\eta_j}{2}\right)
  =
  1 .
\]
Therefore
\[
  \operatorname*{Res}_{\theta'=\theta+\ii\pi}
  P_{n+2}(\theta',\theta,\eta_1,\ldots,\eta_n)
  =
  2P_n(\eta_1,\ldots,\eta_n).
\]
If \(n=2k+1\), then \(F_n^\Sigma=v\ii^kP_n\), while
\(F_{n+2}^\Sigma=v\ii^{k+1}P_{n+2}\).  Hence
\[
  -\ii\,\operatorname*{Res} F_{n+2}^\Sigma
  =
  -\ii\,2v\ii^{k+1}P_n
  =
  2v\ii^kP_n
  =
  2F_n^\Sigma .
\]
Since \(n\) is odd, \(1-(-1)^n=2\), giving the claimed kinematic-pole
equation.

\begin{proposition}[Separated Euclidean spectral series for the Ising branch fields]
\label{prop:ising-spin-separated-euclidean-series}
Use the free Majorana one-particle normalization of
Proposition~\ref{prop:majorana-energy-two-particle-reconstruction}, and
consider Euclidean time separation \(r>0\).  Define
\begin{equation}
  I_m(r)
  =
  \int_{\R}\frac{\dd\theta}{2\pi}\,e^{-mr\cosh\theta}
  =
  \frac{1}{\pi}K_0(mr).
  \label{eq:ising-one-particle-euclidean-majorant}
\end{equation}
The even spin-field form factors
\eqref{eq:ising-spin-even-form-factors} determine an absolutely convergent
separated-point Euclidean spectral series
\begin{equation}
  G_{\sigma_+}(r)
  =
  \sum_{k=0}^{\infty}\frac{1}{(2k)!}
  \int_{\R^{2k}}\prod_{j=1}^{2k}\frac{\dd\theta_j}{2\pi}\,
  \left|F_{2k}^{\sigma_+}(\theta_1,\ldots,\theta_{2k})\right|^2
  e^{-mr\sum_{j=1}^{2k}\cosh\theta_j}.
  \label{eq:ising-spin-even-spectral-series}
\end{equation}
For every \(r>0\),
\begin{equation}
  |G_{\sigma_+}(r)|
  \le
  |\bar\sigma|^2\cosh I_m(r).
  \label{eq:ising-spin-even-majorant}
\end{equation}
After removing the \(k=0\) term, the connected series is bounded by
\(|\bar\sigma|^2(\cosh I_m(r)-1)\).

The odd product family \eqref{eq:ising-order-form-factors} similarly gives
\begin{equation}
  G_{\Sigma}(r)
  =
  \sum_{k=0}^{\infty}\frac{1}{(2k+1)!}
  \int_{\R^{2k+1}}\prod_{j=1}^{2k+1}\frac{\dd\theta_j}{2\pi}\,
  \left|F_{2k+1}^{\Sigma}(\theta_1,\ldots,\theta_{2k+1})\right|^2
  e^{-mr\sum_{j=1}^{2k+1}\cosh\theta_j},
  \label{eq:ising-spin-odd-spectral-series}
\end{equation}
with
\begin{equation}
  |G_{\Sigma}(r)|
  \le
  |v|^2\sinh I_m(r).
  \label{eq:ising-spin-odd-majorant}
\end{equation}
Both series converge uniformly with all \(r\)-derivatives on compact subsets
of \((0,\infty)\).  Hence they define smooth separated Euclidean two-point
functions on the temporal ray; Euclidean covariance then gives the
corresponding separated radial functions.  This proposition is a
separated-point reconstruction statement for this free model.  The full
local-reconstruction problem for a general factorizing \(S\)-matrix is the
open problem stated below.
\end{proposition}

\begin{proof}
The finite-particle identity resolution gives
\eqref{eq:ising-spin-even-spectral-series} and
\eqref{eq:ising-spin-odd-spectral-series} exactly as in the energy-density
derivation, with the factorial compensating for integration over labelled
rapidity variables of identical Majorana particles.  For real rapidities,
\[
  \left|\tanh\!\left(\frac{\theta_i-\theta_j}{2}\right)\right|
  \le 1 .
\]
Therefore
\begin{equation}
  \left|F_{2k}^{\sigma_+}(\theta_1,\ldots,\theta_{2k})\right|^2
  \le
  |\bar\sigma|^2,
  \qquad
  \left|F_{2k+1}^{\Sigma}(\theta_1,\ldots,\theta_{2k+1})\right|^2
  \le
  |v|^2.
  \label{eq:ising-form-factor-product-majorant}
\end{equation}
The factorized exponential then gives the termwise bounds
\begin{equation}
  \frac{1}{(2k)!}
  \int_{\R^{2k}}\prod_{j=1}^{2k}
  \frac{\dd\theta_j}{2\pi}\,
  |\bar\sigma|^2 e^{-mr\sum_j\cosh\theta_j}
  =
  |\bar\sigma|^2\frac{I_m(r)^{2k}}{(2k)!}
  \label{eq:ising-even-term-majorant}
\end{equation}
and
\begin{equation}
  \frac{1}{(2k+1)!}
  \int_{\R^{2k+1}}\prod_{j=1}^{2k+1}
  \frac{\dd\theta_j}{2\pi}\,
  |v|^2 e^{-mr\sum_j\cosh\theta_j}
  =
  |v|^2\frac{I_m(r)^{2k+1}}{(2k+1)!}.
  \label{eq:ising-odd-term-majorant}
\end{equation}
Summing the even and odd majorants gives
\eqref{eq:ising-spin-even-majorant} and
\eqref{eq:ising-spin-odd-majorant}.  The identity
\eqref{eq:ising-one-particle-euclidean-majorant} follows from the integral
definition of \(K_0\) and evenness in \(\theta\).

Let \(0<r_0<r_1<\infty\).  A derivative of order \(q\) with respect to
\(r\) inserts a finite sum of factors bounded by
\[
  m^q\left(\sum_j\cosh\theta_j\right)^q
  e^{-mr\sum_j\cosh\theta_j}.
\]
For \(r\in[r_0,r_1]\), split
\[
  e^{-mr\sum_j\cosh\theta_j}
  =
  e^{-(mr_0/2)\sum_j\cosh\theta_j}
  e^{-m(r-r_0/2)\sum_j\cosh\theta_j}.
\]
The polynomial factor in \(\cosh\theta_j\) is dominated by a constant
depending on \(q,r_0,m\) times the first exponential, and the remaining
integrals are bounded by the same exponential series with \(r_0/2\) in place
of \(r\).  The Weierstrass test therefore applies to every derivative order
on \([r_0,r_1]\).  The resulting functions are smooth for \(r>0\), and the
spectral construction gives the separated Euclidean two-point functions on
the temporal ray.  Contact terms at coincident points are outside this
separated-point estimate and belong to the extension problem on the
collision diagonal.
\end{proof}

\section{Bound-State Poles and Operator Couplings}

Suppose species \(a\) and \(b\) have a bound state \(c\).  If the
two-body \(S\)-matrix has a physical-strip pole at \(\theta=\ii u_{ab}^c\)
with residue tensor \(\Gamma_{ab}^c:V_c\to V_a\otimes V_b\), then the form
factor has the compatible pole
\[
  -\ii\,\operatorname*{Res}_{\theta_a-\theta_b=\ii u_{ab}^c}
  F^{\mathcal O}_{a b a_1\ldots a_n}
  =
  F^{\mathcal O}_{c a_1\ldots a_n}\,\Gamma_{ab}^c ,
\]
again with the stated tensor-ordering convention.  This equation is the
operator-level statement that a local field couples to the bound state through
the same one-particle channel that appears in the scattering residue.

\section{Form-Factor Reconstruction Boundary}

The structure of Definition~\ref{def:form-factor-datum-local-operator} supplies
the bootstrap input.  Its growth entry is essential: it controls whether the
spectral series
for Wightman functions,
\[
  \langle\Omega|\widehat{\mathcal O}(x)\widehat{\mathcal O}(0)|\Omega\rangle
  =
  \sum_{n=0}^{\infty}\frac{1}{n!}
  \sum_{a_1,\ldots,a_n}
  \int d\mu_n\,
  e^{-\ii x\cdot\sum_jp_j}
  F^{\mathcal O}_{a_1\ldots a_n}(\theta)
  \overline{F^{\mathcal O}_{a_1\ldots a_n}(\theta)}
\]
defines a distribution.  Here \(d\mu_n\) is the Lorentz-invariant
\(n\)-particle rapidity measure with the normalization fixed by the
one-particle Hilbert-space convention.

\paragraph{Status checkpoint: from form factors to fields.}
The preceding equations are necessary matrix-element identities for a local
or semi-local field in a factorizing theory.  They become a local-field
construction only after five analytic steps have been supplied in the same
normalization.  First, the finite-particle vectors must sit in a positive
Hilbert space with a common invariant domain on which the smeared operators
are defined and closable.  Second, the spectral sums must converge as
tempered Wightman distributions, with separated Euclidean functions for
\(r>0\) serving only as one part of that construction; this includes control
of the extension to collision diagonals and of contact terms.  Third, the
resulting distributions must satisfy covariance,
the spectrum condition, and positivity.  Fourth, the commutator or
semi-locality relation must vanish on spacelike test-function supports as an
operator statement on the common domain.  Fifth, the finite-particle states
used in the series must be complete in the sector claimed.  The Ising
Majorana examples above meet these requirements because the underlying free
field already supplies the Hilbert space and local fields, and the product
bounds prove separated Euclidean convergence for the displayed branch
families.  For an interacting scalar factorizing model, the same list is a
set of additional analytic hypotheses beyond Watson, cyclicity, and residue
equations.

\begin{openproblem}[Local reconstruction from form factors]
\label{op:form-factor-local-reconstruction}
Given a factorized \(S\)-matrix and a family of form-factor data satisfying
the displayed equations and explicit growth bounds, construct the Wightman
domain, prove convergence of the spectral series as distributions, verify
local commutativity, and identify a dense set of local operators.  This is
the operator-level completion of the scattering bootstrap.
\end{openproblem}
