\chapter{Factorized Scattering and Integrability}
\label{ch:integrable-factorized-scattering}

\ConventionsLorentzian

This volume begins with massive two-dimensional QFT because that is the place
where integrability has a sharp scattering-theoretic meaning.  Let the theory
have a vacuum representation on \(\mathbb M^2\), a mass gap above the vacuum,
and isolated one-particle species labelled by \(a,b,\ldots\), with masses
\(m_a>0\).  A one-particle momentum is parametrized by rapidity
\[
  p_a^\mu(\theta)
  =
  (m_a\cosh\theta,\ m_a\sinh\theta),
  \qquad \theta\in\R .
\]
The Lorentz group acts by translating \(\theta\).  Hence a two-particle
invariant depends on the rapidity difference \(\theta_{12}=\theta_1-\theta_2\).

\begin{definition}[Rapidity chart and ordered chambers]
\label{def:integrable-rapidity-chart-ordered-chambers}
For a species \(a\), the rapidity chart is the map
\[
  \theta\longmapsto p_a(\theta)=m_a(\cosh\theta,\sinh\theta)
\]
from \(\R\) to the positive-energy mass shell
\(\{p\in\R^{1,1}:p^2=-m_a^2,\ p^0>0\}\).  The positive two-particle
Mandelstam invariant is
\[
  s_{ab}(\theta_1-\theta_2)
  :=
  -\bigl(p_a(\theta_1)+p_b(\theta_2)\bigr)^2 .
\]
For \(N\) particles, an ordered rapidity chamber is a connected component of
\[
  \{(\theta_1,\ldots,\theta_N)\in\R^N:
    \theta_i\ne\theta_j\ \text{for }i\ne j\}.
\]
The fundamental chamber used below is
\(\theta_1>\theta_2>\cdots>\theta_N\).  Other chambers are reached by
permuting rapidities and then relating the corresponding scattering bases by
two-body scattering operators.
\end{definition}

\begin{remark}[Two-particle kinematics in rapidity variables]
\label{rem:two-particle-rapidity-kinematics}
In the rapidity chart of
Definition~\ref{def:integrable-rapidity-chart-ordered-chambers},
\[
  s_{ab}(\theta)
  =
  m_a^2+m_b^2+2m_am_b\cosh\theta,
  \qquad \theta=\theta_1-\theta_2 .
\]
The Lorentz-invariant two-body kinematics of massive particles in
\(1+1\) dimensions is therefore encoded by the rapidity difference
\(\theta\), together with the species labels.
\end{remark}

With mostly-plus metric,
\[
  p_a(\theta_1)\cdot p_b(\theta_2)
  =
  -m_am_b\cosh\theta_1\cosh\theta_2
  +m_am_b\sinh\theta_1\sinh\theta_2
  =
  -m_am_b\cosh(\theta_1-\theta_2).
\]
Since \(p_a^2=-m_a^2\) and \(p_b^2=-m_b^2\),
\[
  -\bigl(p_a+p_b\bigr)^2
  =
  m_a^2+m_b^2+2m_am_b\cosh(\theta_1-\theta_2).
\]
A proper orthochronous Lorentz boost of rapidity \(\lambda\) sends
\(\theta_i\mapsto\theta_i+\lambda\), so \(\theta_1-\theta_2\) is invariant.
Conversely, once the masses and the total invariant mass are fixed,
\(\cosh(\theta_1-\theta_2)\) is fixed; the sign of
\(\theta_1-\theta_2\) records the ordering of the two momenta.

\section{Reconstruction Spine and Status Labels}
\label{sec:integrable-reconstruction-spine}

The exact data in this volume do not all have the same logical status.  A
two-body \(S\)-matrix satisfying unitarity, crossing, and Yang--Baxter is
on-shell scattering data.  A TBA equation is a thermodynamic or mirror-channel
functional equation after Bethe--Yang quantization has been accepted.  A
form-factor solution is a family of candidate local-operator matrix elements.
None of these, by itself, is a constructed local quantum field theory.  The
volume therefore uses a common dependency spine with two intertwined chains.
The Hilbert-space chain runs from a local massive QFT with higher-spin charges
to elastic factorized scattering, then to \(S\)-Fock/ZF data, wedge-local
fields, local algebras, and finally local fields and Wightman functions.  The
thermodynamic chain runs from Bethe--Yang finite-volume quantization to TBA
pseudoenergies, mirror finite-size data, and GHD currents.
The arrows in these chains are not equivalent.  Some are theorems under
standard Hilbert-space hypotheses; some are conditional theorems after
analytic and nuclearity assumptions are imposed; some are formal constructions
whose output still lacks positivity or locality; some are numerical or
finite-regulator evidence; and some are open problems.

\begin{definition}[Status labels for integrable-QFT bridges]
\label{def:integrable-reconstruction-status-labels}
For the rest of this volume, a bridge between two layers of exact data carries
one of the following labels.
\begin{enumerate}[label=\textnormal{(\roman*)}]
\item \emph{Theorem}: the stated hypotheses construct the target object in a
  Hilbert space or algebra of local observables and prove the claimed property.
\item \emph{Conditional theorem}: the proof is valid once explicitly stated
  analytic, domain, nuclearity, thermodynamic-limit, or state-space hypotheses
  are supplied.
\item \emph{Formal construction}: the formula produces algebraic or
  distributional candidates, but positivity, locality, convergence, or domain
  control is not yet part of the statement.
\item \emph{Numerical or finite-regulator evidence}: a calculation check
  verifies a finite identity, normalization, sign, or approximation window.  It
  is not evidence for continuum reconstruction unless the limiting theorem is
  separately stated.
\item \emph{Open problem}: a named missing implication whose hypotheses and
  target object are known well enough to be stated.
\end{enumerate}
\end{definition}

The main bridges in this volume should be read with the following status map.
The entries are deliberately conservative; upgrading a row requires proving the
missing analytic statement, not adding another finite identity.
\begin{center}
\small
\begin{tabular}{@{}>{\raggedright\arraybackslash}p{0.23\textwidth}
  >{\raggedright\arraybackslash}p{0.18\textwidth}
  >{\raggedright\arraybackslash}p{0.49\textwidth}@{}}
\textbf{Bridge} & \textbf{Status} & \textbf{Load-bearing evidence} \\
\hline
local QFT to factorized scattering
& conditional theorem
& Haag--Ruelle scattering, mass gap, separating higher-spin charges,
regular distribution kernels, and control of threshold and bound-state
singularities. \\
factorized \(S\) to \(S\)-Fock/ZF algebra
& theorem at the algebraic Hilbert-space level
& Unitarity gives positive chamber inner products; Yang--Baxter gives
path-independent chamber continuation and associative exchange. \\
\(S\)-Fock data to wedge-local fields
& conditional theorem
& Regularity, crossing, and analytic bounds strong enough to define
polarization-free generators and wedge-local operator domains. \\
wedge-local fields to local algebras
& conditional theorem
& Nontriviality of double-cone intersections, usually via modular nuclearity
or related phase-space bounds. \\
form-factor equations to local fields
& formal construction unless convergence and locality are proved
& Watson, cyclicity, and residue equations plus growth bounds; the
missing analytic work is convergence of spectral sums, positivity, domains,
and locality of reconstructed fields. \\
Bethe--Yang data to ground-state TBA
& conditional theorem
& Finite-volume quantization with controlled errors, density convergence,
convex free-energy functional, and fixed normalization of the mirror channel. \\
ground-state TBA to excited-state or mirror data
& conditional or formal, model by model
& Specified singularity motion, contour deformation, source terms, and a check
that no untracked pole crosses the continuation path. \\
TBA densities to GHD
& conditional hydrodynamic limit
& Local generalized Gibbs ensembles, dressing invertibility, Euler-scale
limit, clustering/mixing, and absence or controlled size of integrability
breaking.
\end{tabular}
\end{center}

Two models serve as reference points for this status discipline.  In the
massive Ising model, the scattering datum is the scalar \(S=-1\), the
\(S\)-Fock space is the antisymmetric free-fermion Fock space, and the local
energy field is a Wick polynomial of the Majorana field.  The local QFT and
its Wightman functions are therefore constructed by the free-field theorem;
the form-factor expansion for separated Euclidean correlators is a theorem
after the usual Fock-space domain and convergence estimates.  The spin and
disorder fields require the additional order/disorder semi-locality structure
developed later in the volume, but their status is not inferred merely from
the equation \(S=-1\).

For a regular scalar factorizing model such as sinh-Gordon, the exact
two-body scattering function gives an \(S\)-symmetric Fock space and
wedge-local polarization-free generators under the standard analyticity and
crossing bounds.  The existence of nontrivial double-cone algebras is then a
conditional theorem once the relevant modular nuclearity estimate is proved.
Pointlike local fields obtained from form-factor series, and completeness of a
declared family of such fields, remain separate analytic tasks.  TBA
predictions for the same model are therefore thermodynamic consequences of
the scattering specification under Bethe--Yang and mirror-channel hypotheses;
they are not, by themselves, a proof that the local double-cone net has been
constructed.

This status map is also how calculation companions are used.  A finite
Yang--Baxter, residue, TBA-kernel, or form-factor check verifies the stated
finite algebra in its row of the map.  It does not upgrade a formal
form-factor solution to a local field, nor a TBA equation to a theorem about
finite-size spectra, unless the missing Hilbert-space, convergence, and
thermodynamic-limit assumptions have been supplied at the point of use.

\begin{controlledapproximation}[Wedge-local reconstruction budget from an
\(S\)-matrix]
\label{ca:integrable-wedge-local-reconstruction-budget}
Let \(S(\theta)\) be an elastic factorizing scattering function satisfying
the algebraic conditions in
Definition~\ref{def:elastic-factorized-scattering-system}, and assume the
regularity and crossing bounds needed to define an \(S\)-symmetric Fock space
\(\Hilb_S\), Zamolodchikov--Faddeev operators, and wedge-local
polarization-free generators on a common invariant domain.  Denote the
resulting right-wedge algebra by \(\mathfrak A_S(W_R)\), its translated
wedges by \(\mathfrak A_S(W_R+x)\), and the opposite wedge algebra by
\(\mathfrak A_S(W_L+y)\).  For a double cone written as an intersection of
opposite wedges,
\[
  \mathcal O_{x,y}=(W_R+x)\cap(W_L+y),
  \qquad x\prec y ,
\]
the local-algebra candidate is
\begin{equation}
  \mathfrak A_S(\mathcal O_{x,y})
  :=
  \mathfrak A_S(W_R+x)\cap\mathfrak A_S(W_L+y).
  \label{eq:integrable-double-cone-local-intersection}
\end{equation}
This intersection is the first genuinely local object in the reconstruction
chain.  The \(S\)-Fock space and ZF exchange algebra alone do not prove that
\eqref{eq:integrable-double-cone-local-intersection} contains any nontrivial
observable.

The phase-space estimate used to promote wedge-local data to local algebras
is a modular nuclearity statement.  In the notation of the AQFT nuclearity
discussion in
Chapter~\ref{ch:volume-iv-superselection-sectors-locality-properties}, the
relevant map is represented by
\begin{equation}
  \Xi_{x,y}(A)
  =
  \Delta_{W_R}^{1/4}\,U(y-x)\,A\Omega,
  \qquad
  A\in\mathfrak A_S(W_R),\quad \|A\|\le 1 ,
  \label{eq:integrable-modular-nuclearity-map}
\end{equation}
with the precise translation vector and wedge chosen so that the translated
right wedge lies inside the opposite wedge defining
\(\mathcal O_{x,y}\).  A nuclear norm bound on \(\Xi_{x,y}\), uniform in the
declared double-cone window, is a conditional-theorem input for nontrivial
local intersections, Reeh--Schlieder-type cyclicity on the local algebra, and
the construction of a local net.  Unitarity, Yang--Baxter, and the finite
chamber-extension algebra supply the on-shell \(S\)-Fock coordinate; the
phase-space estimate is an additional local-observable input.

Consequently an observable reconstructed from factorized scattering carries
the residual budget
\begin{equation}
\begin{aligned}
  \mathcal W_{\rm loc}[\varphi]
  -
  \mathcal W_{\rm lead}^{S}[\varphi]
  &=
  R_{\rm pfg}[\varphi]
  +R_{\rm nuc}[\varphi]
  +R_{\rm int}[\varphi]
  +R_{\rm ff}[\varphi]  \\
  &\qquad
  +R_{\rm dom}[\varphi]
  +R_{\rm comp}[\varphi] .
\end{aligned}
  \label{eq:integrable-local-reconstruction-residual-budget}
\end{equation}
Here \(\mathcal W_{\rm lead}^{S}\) is the candidate observable coordinate
computed from the retained \(S\)-Fock or form-factor data; \(R_{\rm pfg}\)
records failures of the wedge-local generator construction, analytic
regularity, or common domains; \(R_{\rm nuc}\) records the modular nuclearity
or phase-space estimate; \(R_{\rm int}\) records nontriviality, isotony, and
covariance of the double-cone intersections; \(R_{\rm ff}\) records
form-factor convergence and collision/contact extensions; \(R_{\rm dom}\)
records operator-domain, positivity, and Wightman/OS distribution issues; and
\(R_{\rm comp}\) records scattering-state and local-field completeness.
If each term is controlled on the same test-function window by
\[
  |R_\kappa[\varphi]|\le\varepsilon_\kappa\|\varphi\|_{\mathcal T},
  \qquad
  \kappa\in
  \{{\rm pfg},{\rm nuc},{\rm int},{\rm ff},{\rm dom},{\rm comp}\},
\]
then
\begin{equation}
  \left|
    \mathcal W_{\rm loc}[\varphi]
    -
    \mathcal W_{\rm lead}^{S}[\varphi]
  \right|
  \le
  \left(
    \varepsilon_{\rm pfg}
    +\varepsilon_{\rm nuc}
    +\varepsilon_{\rm int}
    +\varepsilon_{\rm ff}
    +\varepsilon_{\rm dom}
    +\varepsilon_{\rm comp}
  \right)\|\varphi\|_{\mathcal T}.
  \label{eq:integrable-local-reconstruction-residual-bound}
\end{equation}
This is a reconstruction budget, not another exact scattering identity.  The
rows of the status table enter at different places: Yang--Baxter and
unitarity build the \(S\)-Fock coordinate; regular crossing and wedge-domain
control build wedge generators; modular nuclearity makes
\(\mathfrak A_S(\mathcal O)\) nontrivial; form-factor estimates construct
pointlike or smeared local fields affiliated with that local algebra; and
completeness identifies the resulting local observables with the intended
QFT.

In the massive Ising model the same budget collapses because the local net is
already supplied by the free Majorana field: the wedge and local algebras,
domains, positivity, and energy-density form-factor convergence are inherited
from the free-field construction.  For a regular interacting scalar
factorizing model, by contrast, exact \(S\)-matrix algebra and strong
form-factor evidence still leave the nuclearity, local-intersection, domain,
and completeness entries as separate analytic claims.  This distinction is
the physics of reconstruction: it decides whether the exact on-shell data
have produced local observables or only a beautifully consistent scattering
calculus.
\end{controlledapproximation}

\begin{controlledapproximation}[End-to-end observable reconstruction map]
\label{ca:integrable-end-to-end-reconstruction-route-ledger}
For a concrete integrable model, an ``exact solution'' in this volume means a
map from exact data to a specified observable through stated checkpoints, not a
single formula.  The route datum is the ordered reference map
\[
  \mathsf R
  =
  \bigl(
  \mathsf S,\mathsf H,\mathsf W,\mathsf O,
  \mathsf F,\mathsf T,\mathsf P
  \bigr),
\]
where \(\mathsf S\) is the on-shell scattering or Bethe datum,
\(\mathsf H\) is the Hilbert-space representation, \(\mathsf W\) is the
wedge-local construction, \(\mathsf O\) is the nontrivial double-cone local
observable algebra, \(\mathsf F\) is the pointlike or smeared form-factor
operator package with domains and convergence, \(\mathsf T\) is the
thermodynamic or mirror-state limit, and \(\mathsf P\) is the physical
projection to the observable being quoted.  A local-observable claim requires
all checkpoints on the path actually used by the observable.  A thermodynamic
claim may bypass point fields only if its own finite-volume, Bethe--Yang,
mirror, and state-limit hypotheses replace them.  Thus \(\mathsf R\) records
the physical output, the strongest supplied evidence at each checkpoint, and the
precise places where an exact formula has not yet become a local,
thermodynamic, hydrodynamic, or transport observable.

The following four observable maps fix how later chapters should be read.
\begin{enumerate}[label=\textnormal{(\alph*)}]
\item \emph{Ising energy.}  The controlled output is the local Euclidean
two-point function.  The free Majorana CAR construction supplies
\(\mathsf H,\mathsf W,\mathsf O\), domains, positivity, and completeness; the
two-particle form-factor formula is then a matrix-element calculation inside an
already constructed local QFT.
\item \emph{Sinh-Gordon wedge.}  The controlled output is a conditional local
net.  A regular scalar \(S(\theta)\) supplies \(S\)-Fock and wedge-local fields
under analyticity and crossing bounds, but \(\mathsf O\) requires modular
nuclearity and point-field claims still require \(\mathsf F\) and completeness
estimates.
\item \emph{Lee--Yang TBA.}  The controlled output is a thermodynamic scaling
function.  The TBA equation gives a controlled thermodynamic coordinate only
after finite-volume Bethe--Yang errors, contour and branch data, and state-limit
hypotheses are supplied; equality of endpoint central charges is not a local
trace-operator reconstruction.
\item \emph{GHD cell.}  The controlled output is an Euler-scale density or
current.  Root-density continuity and dressing give the Euler candidate, but
microscopic density and current operators, local-equilibrium formation,
gradients, diffusion, and integrability-breaking residuals remain part of
\(\mathsf P\).  Ballistic transport coefficients add a real-time KMS/Kubo
gate: the dressed-current formula supplies the Drude coordinate only after the
microscopic current and long-time limit have been identified.
\end{enumerate}

Thus the four reference routes have different proof statuses for physical
reasons, not because some have more formulas.  The Ising energy-density route is
theorem-level because locality and domains are inherited before the form-factor
sum is evaluated.  The Sinh-Gordon route is a conditional local-net
construction until the phase-space and local-intersection estimates are in
place, and it is still not a point-field completeness theorem without the
separate form-factor growth and domain input.  The Lee--Yang and GHD routes are
thermodynamic/hydrodynamic routes: they can be excellent QFT predictions, but
their missing gates are finite-volume state limits, microscopic observable
identification, real-time Kubo limits for transport claims, and
local-equilibrium control rather than Yang--Baxter algebra.

If \(E_{\mathsf R}^{\rm obs}\) denotes the error in the target local or
thermodynamic observable along a chosen route, and if every retained residual
has been represented in the same observable norm or topology, the reconstruction
map is used through the absolute budget
\begin{equation}
  |E_{\mathsf R}^{\rm obs}|
  \le
  \varepsilon_{\mathsf S}
  +\varepsilon_{\mathsf H}
  +\varepsilon_{\mathsf W}
  +\varepsilon_{\mathsf O}
  +\varepsilon_{\mathsf F}
  +\varepsilon_{\mathsf T}
  +\varepsilon_{\mathsf P},
  \label{eq:integrable-route-ledger-budget}
\end{equation}
with omitted checkpoints assigned zero only when they are genuinely irrelevant
to the observable's route, not when their analysis has merely been skipped.
Before such an estimate is supplied, \(\varepsilon_{\mathsf S},\ldots,
\varepsilon_{\mathsf P}\) are named residual slots, not numerical bounds; a slot
marks where the physics still owes a construction or limiting argument.  Exact
Yang--Baxter algebra sets the already-defined \(\varepsilon_{\mathsf S}\) to
zero; it says nothing about \(\varepsilon_{\mathsf O}\),
\(\varepsilon_{\mathsf F}\), or \(\varepsilon_{\mathsf P}\).  Likewise an exact
numerical TBA solution may set the retained \(\mathsf T\)-coordinate error to
zero while leaving the local trace-operator or microscopic-current projection
errors untouched.
\end{controlledapproximation}

\section{Factorized Scattering Datum}

\begin{definition}[Elastic factorized scattering system]
\label{def:elastic-factorized-scattering-system}
An elastic factorized scattering system consists of a finite or countable set
of massive species, a positive mass \(m_a\) for each species, and meromorphic
functions
\[
  S_{ab}^{cd}(\theta),
  \qquad a,b,c,d\in I,
\]
defined on a common rapidity-domain cover, satisfying the following algebraic
and analytic conditions.

First, two-particle unitarity is
\[
  \sum_{e,f}
  S_{ab}^{ef}(\theta)\,
  S_{ef}^{cd}(-\theta)
  =
  \delta_a^{c}\delta_b^{d}.
\]
Second, Hermitian analyticity is
\[
  S_{ab}^{cd}(\theta)^\ast
  =
  S_{cd}^{ab}(-\theta^\ast)
\]
on the boundary where both sides are defined.  Third, if
\(\bar a\) denotes the charge-conjugate species, crossing has the form
\[
  S_{ab}^{cd}(\theta)
  =
  C_{a\bar a'}C^{c\bar c'}\,
  S_{\bar c' b}^{\bar a' d}(\ii\pi-\theta),
\]
with the charge-conjugation matrix \(C\) fixed as part of the particle-state
normalization.  Fourth, three-particle consistency is the Yang--Baxter
equation
\[
  S_{12}(\theta_{12})S_{13}(\theta_{13})S_{23}(\theta_{23})
  =
  S_{23}(\theta_{23})S_{13}(\theta_{13})S_{12}(\theta_{12})
\]
as an identity on the tensor product of three one-particle internal spaces.
\end{definition}

The notation \(S_{ij}\) means that the two-particle matrix \(S\) acts on the
\(i\)-th and \(j\)-th tensor factors and as the identity on the remaining
factor.  The Yang--Baxter equation is the associativity condition for
reordering three rapidities by adjacent transpositions.  It is a structural
part of factorized scattering: it identifies the pairwise decompositions of a
single ordered three-particle process.

\begin{definition}[Separating higher-spin family]
\label{def:separating-higher-spin-family}
Let \(\{Q_{s,\lambda}\}_{(s,\lambda)\in\mathcal A}\) be commuting conserved
charges whose one-particle eigenvalues have the form
\[
  q_{s,\lambda,a}(\theta)=\kappa_{s,\lambda,a}\,\ee^{s\theta},
  \qquad s>1.
\]
For fixed particle number \(N\), set \(z_i=\ee^{\theta_i}\).  The family is
\(N\)-separating on a rapidity domain \(U_N\) if the moment map
\[
  \Phi_N\bigl(\{(a_i,z_i)\}_{i=1}^N\bigr)
  =
  \left(
    \sum_{i=1}^N\kappa_{s,\lambda,a_i}z_i^s
  \right)_{(s,\lambda)\in\mathcal A}
\]
is injective on \(U_N/\mathfrak S_N\), outside a specified closed exceptional
set consisting of coincident rapidities and declared threshold or bound-state
singular loci.  The quotient means that only the unordered multiset of
species and rapidities is to be recovered.

A concrete sufficient condition, for a finite species set, is the following.
For each species \(a\) and for each \(r\) needed up to the maximal particle
number under consideration, finite linear combinations of the charges produce
the species-resolved power sums
\[
  p_{r,a}=\sum_{i:\,a_i=a}z_i^r .
\]
Then Newton identities recover the monic polynomial
\[
  \prod_{i:\,a_i=a}(t-z_i)
\]
for each species \(a\), and hence recover the unordered species-resolved
rapidity multiset wherever the roots are distinct.  The definition above is
the hypothesis used below; this sufficient condition is only a convenient
checkable way to obtain it.
\end{definition}

\paragraph{Support test by multiplication.}
Let \(U\subset\R^n\) be open, let \(T\in\mathcal D'(U)\), and let
\(f\in C^\infty(U)\).  If \(fT=0\) and \(f\) has no zero on an open set
\(V\subset U\), then \(T\) vanishes on \(V\).

For any test function \(\varphi\in C_c^\infty(V)\), the quotient
\(\varphi/f\) is smooth with compact support in \(V\), because \(f\) is
nowhere zero on \(V\).  Since \(fT=0\),
\[
  \langle T,\varphi\rangle
  =
  \left\langle T,f\frac{\varphi}{f}\right\rangle
  =
  \left\langle fT,\frac{\varphi}{f}\right\rangle
  =
  0 .
\]
Thus \(T\) pairs to zero with every test function supported in \(V\), which
means \(T|_V=0\).

\begin{proposition}[Factorization from conserved higher-spin charges]
\label{prop:higher-spin-factorization-kinematic}
Assume a massive \(1+1\)-dimensional QFT has Haag--Ruelle asymptotic particle
states, a family of conserved charges
\(\{Q_{s,\lambda}\}_{(s,\lambda)\in\mathcal A}\) as in
Definition~\ref{def:separating-higher-spin-family}, and scattering kernels
which, after smearing by compactly supported wave packets in rapidity
chambers, depend continuously on the wave packets away from threshold and
bound-state singular loci.  Equivalently, the unsmeared kernels are
distributional boundary values whose singular support in the rapidity
variables is contained in those declared loci.

If the scattering operator commutes with every \(Q_{s,\lambda}\), and if the
charge family is \(N\)-separating on the incoming and outgoing \(N\)-particle
domains under consideration, then the \(N\to N\) scattering kernel is
supported, away from the declared singular loci, on the union of permutation
graphs
\[
  \{(\theta'_1,b_1;\ldots;\theta'_N,b_N)
    =
    (\theta_{\sigma(1)},a_{\sigma(1)};\ldots;
     \theta_{\sigma(N)},a_{\sigma(N)})\}.
\]
If the corresponding moment maps for \(M\neq N\) have disjoint generic
images, then generic \(N\to M\) particle production is absent.
\end{proposition}

\begin{proof}
Let \(K_{N\to M}\) be the distribution kernel of the scattering operator
between ordered rapidity chambers, away from the declared singular loci.
Commutation with \(Q_{s,\lambda}\) gives, in the sense of distributions,
\[
  \left(
    \sum_{i=1}^N\kappa_{s,\lambda,a_i}\ee^{s\theta_i}
    -
    \sum_{j=1}^M\kappa_{s,\lambda,b_j}\ee^{s\theta'_j}
  \right)
  K_{N\to M}=0
\]
for every \((s,\lambda)\).  The following support property of distributions
is being used: the support test above implies that
the kernel vanishes on any open set where one of the displayed charge
differences is nonzero.  Therefore the support of the kernel is contained in
the common zero locus of all differences of charge eigenvalues.

For \(M=N\), the \(N\)-separation hypothesis says that this common zero locus,
outside the exceptional set, is precisely the union of permutation graphs
displayed in the statement.  Thus, after smearing with wave packets supported
in a connected component of the regular rapidity chamber, the scattering
operator cannot have support away from elastic permutations.  For \(M\neq N\),
the assumed disjointness of the generic moment-map images leaves no regular
common zero locus, so the production kernel vanishes on the corresponding
open domain.  Threshold and bound-state loci are excluded from the statement
because the distribution kernels may have additional singular components
there; those components must be described after the relevant channel is
resolved.
\end{proof}

\section{Ordered Bases and the Scattering Groupoid}

Factorized scattering is most precise when formulated as a comparison of
ordered bases for the same asymptotic Hilbert-space sector.  Fix a regular
rapidity point with distinct rapidities.  For each chamber
\(\theta_{\sigma(1)}>\cdots>\theta_{\sigma(N)}\), choose the ordered
asymptotic basis in which the creation operators are written in that order.
Crossing a codimension-one wall \(\theta_i=\theta_{i+1}\) is represented by
the two-body operator
\[
  \mathsf T_i(\theta_i-\theta_{i+1})
  =
  S_{i,i+1}(\theta_i-\theta_{i+1})
\]
acting on the adjacent species spaces and reversing the adjacent rapidity
order.  Thus the elementary maps \(\mathsf T_i\) are morphisms in the groupoid
whose objects are rapidity chambers and whose arrows are paths crossing
walls.  A consistent factorized scattering theory requires this groupoid
representation to be path independent on the complement of the declared
threshold and bound-state singular loci.

\begin{proposition}[Scattering-chamber consistency]
\label{prop:scattering-chamber-consistency}
Let \(S_{ab}^{cd}(\theta)\) be an elastic factorized scattering system.  Away
from poles and coincident rapidities, the adjacent chamber maps
\(\mathsf T_i\) define a path-independent representation of the chamber
groupoid if and only if the two-particle unitarity identities and the
Yang--Baxter equation hold.
\end{proposition}

\begin{proof}
The chamber groupoid for ordered points on the line is generated by adjacent
wall crossings.  Its local relations are:
\[
  \mathsf T_i^2=1,\qquad
  \mathsf T_i\mathsf T_j=\mathsf T_j\mathsf T_i\quad (|i-j|>1),
  \qquad
  \mathsf T_i\mathsf T_{i+1}\mathsf T_i
  =
  \mathsf T_{i+1}\mathsf T_i\mathsf T_{i+1}.
\]
The first relation is exactly two-particle unitarity,
\[
  S_{ab}(\theta)S_{ba}(-\theta)=\id,
\]
written in components as the identity in
Definition~\ref{def:elastic-factorized-scattering-system}.  The second
relation holds because disjoint adjacent crossings act on disjoint tensor
factors.  For three adjacent rapidities
\(\theta_1>\theta_2>\theta_3\), the two reduced paths from the chamber
\((123)\) to the chamber \((321)\) give
\[
  S_{12}(\theta_{12})S_{13}(\theta_{13})S_{23}(\theta_{23})
  \quad\text{and}\quad
  S_{23}(\theta_{23})S_{13}(\theta_{13})S_{12}(\theta_{12}),
\]
respectively.  Equality is precisely the Yang--Baxter equation.  These local
relations present the chamber groupoid away from the excluded diagonals and
singular loci, so unitarity plus Yang--Baxter gives path independence.
Conversely, path independence around the two local loops just described gives
unitarity and the Yang--Baxter equation.
\end{proof}

\section{Zamolodchikov--Faddeev Operators}

The factorized scattering system becomes an operator algebra on asymptotic wavefunctions.
For each species \(a\) introduce distributional creation operators
\(Z_a^\dagger(\theta)\) obeying
\begin{equation}
  Z_a^\dagger(\theta_1)Z_b^\dagger(\theta_2)
  =
  S_{ab}^{cd}(\theta_1-\theta_2)\,
  Z_d^\dagger(\theta_2)Z_c^\dagger(\theta_1),
  \qquad \theta_1>\theta_2 .
  \label{eq:zf-creation-exchange}
\end{equation}
Repeated species labels are summed.  These operator-valued distributions are
a compact encoding of ordered asymptotic creation operators and their
chamber-transition functions.  Local fields enter through form factors and
operator reconstruction below.

\paragraph{Associativity of the ZF creation algebra.}
Assume two-particle unitarity.  The exchange rule
\eqref{eq:zf-creation-exchange} gives an associative algebra of ordered
creation symbols, independent of the sequence of adjacent re-orderings, if
and only if the Yang--Baxter equation in
Definition~\ref{def:elastic-factorized-scattering-system} holds.

It is enough to compare the two ways of reordering three strictly ordered
rapidities \(\theta_1>\theta_2>\theta_3\).  Starting from
\[
  Z_a^\dagger(\theta_1)Z_b^\dagger(\theta_2)Z_c^\dagger(\theta_3),
\]
the reduced word which first moves the first rapidity past the second and
then through the third gives the coefficient
\[
  S_{ab}^{a'b'}(\theta_{12})
  S_{a'c}^{a''c'}(\theta_{13})
  S_{b'c'}^{b''c''}(\theta_{23})
\]
multiplying the final ordered product
\[
  Z_{c''}^\dagger(\theta_3)
  Z_{b''}^\dagger(\theta_2)
  Z_{a''}^\dagger(\theta_1).
\]
The other reduced word gives the same final ordered product with coefficient
\[
  S_{bc}^{b'c'}(\theta_{23})
  S_{ac'}^{a'c''}(\theta_{13})
  S_{a'b'}^{a''b''}(\theta_{12}).
\]
Equality of these coefficients for all external species is exactly the
component form of the Yang--Baxter equation.  The inverse adjacent move is
well defined by two-particle unitarity, and disjoint moves commute because
they involve disjoint creation symbols.  Hence the braid relations are
equivalent to associativity and sequence independence of the exchange
algebra.

\begin{definition}[\(S\)-symmetric finite-particle wavefunctions]
\label{def:s-symmetric-finite-particle-wavefunctions}
An \(N\)-particle \(S\)-symmetric wavefunction is a collection of functions
\[
  \psi_{a_1\ldots a_N}(\theta_1,\ldots,\theta_N)
\]
on rapidity chambers, square-integrable with respect to
\(\dd\theta_1\cdots\dd\theta_N\) on compactly supported wave packets, whose
boundary values across an adjacent chamber wall obey
\[
  \psi_{\ldots a_i a_{i+1}\ldots}
    (\ldots,\theta_i,\theta_{i+1},\ldots)
  =
  S_{a_i a_{i+1}}^{bc}(\theta_i-\theta_{i+1})\,
  \psi_{\ldots c b\ldots}
    (\ldots,\theta_{i+1},\theta_i,\ldots).
\]
The finite-particle \(S\)-Fock space is obtained by imposing the
\(L^2\) positive inner product on compactly supported wavefunctions in one
fundamental chamber, extending them to all chambers by the displayed
relations, and completing the resulting pre-Hilbert space.
\end{definition}

\paragraph{Well-defined chamber extension.}
If the two-body scattering system satisfies unitarity and the Yang--Baxter
equation, then
the chamber relations in
Definition~\ref{def:s-symmetric-finite-particle-wavefunctions} determine the
values of an \(S\)-symmetric wavefunction in every chamber from its value in
one fundamental chamber, independently of the chosen adjacent-transposition
path.

Crossing a wall that exchanges adjacent rapidities applies the adjoint
two-body scattering matrix to the corresponding tensor factors.  Two-particle
unitarity makes this wall-crossing map invertible and implements
\(\mathsf T_i^2=1\).  Wall crossings involving disjoint adjacent pairs act on
disjoint tensor factors, hence commute.  The only remaining ambiguity in a
path through adjacent chambers is the braid move
\(\mathsf T_i\mathsf T_{i+1}\mathsf T_i=
\mathsf T_{i+1}\mathsf T_i\mathsf T_{i+1}\), and this is exactly the
Yang--Baxter equation for the three involved tensor factors.  Therefore every
word representing the same permutation gives the same continuation from the
fundamental chamber.

\section{Local Operators and Form Factors}

Let \(\mathcal O(x)\) be a local operator whose matrix elements between the
vacuum and ordered out-states exist as distributions.  Its \(n\)-particle form
factor is
\[
  F^{\mathcal O}_{a_1\ldots a_n}(\theta_1,\ldots,\theta_n)
  =
  \bra{\vac}\mathcal O(0)
  Z_{a_1}^\dagger(\theta_1)\cdots Z_{a_n}^\dagger(\theta_n)\ket{\vac},
  \qquad \theta_1>\cdots>\theta_n .
\]
Changing the ordered rapidity chamber changes the scattering basis by the
two-body \(S\)-matrix.

\begin{remark}[Watson exchange from ordered asymptotic bases]
\label{rem:watson-exchange-ordered-bases}
For adjacent rapidities in a regular chamber, the form-factor boundary values
obey
\[
  F^{\mathcal O}_{\ldots a_i a_{i+1}\ldots}(\ldots,\theta_i,\theta_{i+1},\ldots)
  =
  S_{a_i a_{i+1}}^{bc}(\theta_i-\theta_{i+1})\,
  F^{\mathcal O}_{\ldots c b\ldots}(\ldots,\theta_{i+1},\theta_i,\ldots),
\]
with summation over \(b,c\).  This statement uses only the factorized
scattering normalization of ordered asymptotic bases.  The role of locality
of \(\mathcal O\) begins with the cyclicity and semi-locality conditions
developed in
Chapter~\ref{ch:integrable-form-factor-bootstrap-local-operators}.
\end{remark}

Take the defining matrix element in the chamber
\(\theta_1>\cdots>\theta_i>\theta_{i+1}>\cdots>\theta_n\).  Applying the ZF
exchange relation \eqref{eq:zf-creation-exchange} to the adjacent pair gives
\[
  Z_{a_i}^\dagger(\theta_i)Z_{a_{i+1}}^\dagger(\theta_{i+1})
  =
  S_{a_i a_{i+1}}^{bc}(\theta_i-\theta_{i+1})\,
  Z_c^\dagger(\theta_{i+1})Z_b^\dagger(\theta_i).
\]
Insert this identity between \(\bra{\vac}\mathcal O(0)\) and \(\ket{\vac}\),
leaving all other ordered creation operators fixed.  The resulting matrix
element on the right is exactly the boundary value in the neighboring chamber
with the adjacent rapidities interchanged.  This gives the displayed Watson
equation.  No commutation of \(\mathcal O\) through the ZF operator has been
used.

Together with Watson exchange, form-factor bootstrap uses crossing,
cyclicity, and residue equations at bound-state poles.  The exchange equation
is a statement about boundary values in neighboring rapidity chambers and the
normalization of ordered scattering states; locality of \(\mathcal O\) enters
only in those additional cyclicity and semi-locality conditions.
The operator reconstruction problem is to determine which solutions of these
functional equations come from genuine local operators in a Hilbert-space
QFT.  The theorem-boundary data include the local operator domain,
positivity, convergence of form-factor expansions, and clustering.

\begin{openproblem}[Completeness of form-factor reconstruction]
  \label{op:form-factor-reconstruction-completeness}
Give intrinsic hypotheses on a factorized scattering system and a family of
form-factor solutions that imply the existence of a local QFT with those
asymptotic particles, those local operators, and Wightman functions obtained
by convergent form-factor sums.  Existing constructive approaches prove this
for important classes of two-dimensional models; the monograph will develop
the exact hypotheses and examples in later chapters of this volume.
\end{openproblem}
