\chapter{Nested TBA, Baxter Relations, and Separation Variables}
\label{ch:integrable-nested-tba-tq-separation-variables}

\ConventionsLorentzian

Diagonal TBA converts scalar Bethe--Yang equations into density and
pseudoenergy equations.  Non-diagonal scattering adds auxiliary roots.  The
thermodynamic limit then contains physical densities and magnonic auxiliary
densities.  This chapter develops the finite-dimensional functional
relations that control those auxiliary roots: Baxter \(TQ\) equations,
\(QQ\)-systems, Hirota \(T\)-systems, and separation of variables.  These are
algebraic objects first; their use in a continuum QFT requires the separate
large-volume and thermodynamic hypotheses stated below.

\section{Nested Densities}

Start from logarithmic nested equations of the form
\[
  Lp_a(\theta_j)
  +
  \sum_{b,k}\delta_{ab}(\theta_j-\theta_k)
  +
  \sum_{\alpha,r}\chi_{a r}(\theta_j-u_\alpha^{(r)})
  =
  2\pi I_j ,
\]
together with auxiliary equations
\[
  \sum_{b,k}\chi_{r b}(u_\alpha^{(r)}-\theta_k)
  +
  \sum_{s,\beta}\psi_{rs}(u_\alpha^{(r)}-u_\beta^{(s)})
  =
  2\pi J_\alpha^{(r)} .
\]
The labels \(a,b\) enumerate physical particle species; \(r,s\) enumerate
auxiliary Dynkin nodes or nesting levels.  The functions \(\delta,\chi,\psi\)
are chosen branches of logarithms of the scalar and matrix-scattering
eigenvalue factors.

In the thermodynamic limit introduce physical densities
\(\rho_a,\rho_a^h\) and auxiliary densities
\(\sigma_r,\sigma_r^h\).  Differentiating the logarithmic equations gives
\begin{align}
  \rho_a+\rho_a^h
  &=
  \frac{1}{2\pi}\frac{\dd p_a}{\dd\theta}
  +\sum_b K_{ab}*\rho_b
  +\sum_r K_{a r}*\sigma_r ,
  \label{eq:nested-tba-physical-density}\\
  \sigma_r+\sigma_r^h
  &=
  \sum_a K_{r a}*\rho_a
  +\sum_s K_{rs}*\sigma_s .
  \label{eq:nested-tba-aux-density}
\end{align}
Here
\[
  (K*f)(x)=\int_\R K(x-y)f(y)\,\dd y,
\]
and each kernel is \(1/(2\pi)\) times the derivative of the corresponding
phase.  Auxiliary roots carry no direct one-particle energy in the examples
under discussion; they affect the free energy through the constraints.

\begin{proposition}[Nested TBA variational equations]
\label{prop:nested-tba-variational}
Assume the density equations
\eqref{eq:nested-tba-physical-density}--\eqref{eq:nested-tba-aux-density}
have a thermodynamic limit with positive total densities, and use the
fermionic occupancy entropy for each physical and auxiliary root species.
Define pseudoenergies
\[
  \epsilon_a=\log\frac{\rho_a^h}{\rho_a},
  \qquad
  \eta_r=\log\frac{\sigma_r^h}{\sigma_r}.
\]
Then stationarity of the constrained free energy at inverse temperature \(R\)
gives
\begin{align}
  \epsilon_a
  &=
  R E_a
  -\sum_b K_{ba}*\log(1+\ee^{-\epsilon_b})
  -\sum_r K_{r a}*\log(1+\ee^{-\eta_r}),
  \label{eq:nested-tba-epsilon}\\
  \eta_r
  &=
  -\sum_a K_{a r}*\log(1+\ee^{-\epsilon_a})
  -\sum_s K_{s r}*\log(1+\ee^{-\eta_s}).
  \label{eq:nested-tba-eta}
\end{align}
\end{proposition}

\begin{proof}
Let \(A\) denote either a physical label \(a\) or an auxiliary label \(r\).
Write \(n_A\) for the occupied density, \(n_A^h\) for the hole density, and
\(n_A^t=n_A+n_A^h\) for the total density.  The density equations have the
uniform form
\begin{equation}
  n_A^t=d_A+\sum_B K_{AB}*n_B,
  \qquad
  d_a=\frac1{2\pi}\frac{\dd p_a}{\dd\theta},
  \qquad
  d_r=0 ,
  \label{eq:nested-tba-unified-density-proof}
\end{equation}
where the energy is \(E_a\) on physical labels and \(0\) on auxiliary labels.
The entropy density of one fermionic occupancy species is
\[
  s_A=n_A^t\log n_A^t-n_A\log n_A-n_A^h\log n_A^h,
\]
so, using \(n_A^h=n_A^t-n_A\),
\[
  \delta s
  =
  \sum_A\int \left[
    \log\frac{n_A^h}{n_A}\,\delta n_A
    +
    \log\frac{n_A^t}{n_A^h}\,\delta n_A^t
  \right].
\]
The dimensionless free-energy functional is
\[
  \mathcal F_R[n]
  =
  \sum_a\int R E_a\,\rho_a\,\dd\theta-\sum_A\int s_A\,\dd\theta .
\]
Varying \eqref{eq:nested-tba-unified-density-proof} gives
\[
  \delta n_A^t=\sum_BK_{AB}*\delta n_B .
\]
For test variations of compact support the convolution pairing satisfies
\[
  \int L_A(\theta)(K_{AB}*\delta n_B)(\theta)\,\dd\theta
  =
  \int (K_{AB}^{\dagger}*L_A)(u)\,\delta n_B(u)\,\dd u,
  \qquad
  K_{AB}^{\dagger}(x)=K_{AB}(-x).
\]
With
\[
  \epsilon_A=\log\frac{n_A^h}{n_A},
  \qquad
  L_A=\log\frac{n_A^t}{n_A^h}=\log(1+\ee^{-\epsilon_A}),
\]
stationarity of \(\mathcal F_R\) therefore gives, for every label \(B\),
\begin{equation}
  0=
  R E_B-\epsilon_B-\sum_A K_{AB}^{\dagger}*L_A ,
  \qquad E_r=0 .
  \label{eq:nested-tba-euler-lagrange-proof}
\end{equation}
Writing the adjoint-transposed kernels as \(K_{BA}\), separating physical and
auxiliary labels, and substituting the definitions of
\(\epsilon_a,\eta_r\) gives exactly
\eqref{eq:nested-tba-epsilon}--\eqref{eq:nested-tba-eta}.
\end{proof}

\section{Constant A2 Magnonic Example}

The ultraviolet limit of many nested systems reduces to constant equations
for the plateau values \(Y_A=\ee^{\epsilon_A}\).  A minimal \(A_2\) magnonic
example has two auxiliary nodes and incidence matrix
\[
  I_{rs}=
  \begin{pmatrix}
    0&1\\
    1&0
  \end{pmatrix}.
\]
The constant Y-system is
\[
  Y_1^2=1+Y_2,\qquad
  Y_2^2=1+Y_1 .
\]
The symmetric solution has \(Y_1=Y_2=\phi_{\rm g}\), where
\[
  \phi_{\rm g}=\frac{1+\sqrt5}{2},\qquad
  \phi_{\rm g}^2=1+\phi_{\rm g}.
\]
This finite calculation is the algebraic core behind Rogers-dilogarithm
central-charge evaluations.  To turn it into a central-charge theorem for a
specific QFT one must also justify the scaling limit, the kernel reduction to
the constant Y-system, and the absence or subtraction of bulk-energy terms.

\section{An SU(3) Nested Density System}

The finite \(SU(3)\) equations
\eqref{eq:su3-first-level}--\eqref{eq:su3-second-level} give a useful
nontrivial density example.  Define
\[
  \varphi_\alpha(u)
  =
  \ii\log\frac{u+\ii\alpha}{u-\ii\alpha},
  \qquad
  a_\alpha(u)=\frac{1}{2\pi}\varphi_\alpha'(u)
  =
  \frac{1}{2\pi}\frac{2\alpha}{u^2+\alpha^2}.
\]
The logarithmic branches are fixed continuously along the real root lines.
For first-level roots \(u_j\) and second-level roots \(v_\alpha\), the
counting functions may be chosen as
\[
\begin{aligned}
  Z_u(u)
  &=
  L\varphi_{1/2}(u)
  -\sum_k\varphi_1(u-u_k)
  +\sum_\beta\varphi_{1/2}(u-v_\beta),\\
  Z_v(v)
  &=
  -\sum_\beta\varphi_1(v-v_\beta)
  +\sum_j\varphi_{1/2}(v-u_j).
\end{aligned}
\]
Differentiating and passing to densities gives
\begin{align}
  \rho+\rho^h
  &=
  a_{1/2}-a_1*\rho+a_{1/2}*\sigma ,
  \label{eq:su3-nested-density-first}\\
  \sigma+\sigma^h
  &=
  a_{1/2}*\rho-a_1*\sigma .
  \label{eq:su3-nested-density-second}
\end{align}
The signs are not arbitrary: they are the logarithmic derivatives of the
same-level scattering factors and adjacent-node inverse factors in
\eqref{eq:su3-first-level}--\eqref{eq:su3-second-level}.  If
\(\widehat f(\omega)=\int_\R \ee^{\ii\omega u}f(u)\,\dd u\), then
\[
  \widehat a_\alpha(\omega)=\ee^{-\alpha|\omega|},
\]
so the coupled linear density constraints are diagonalized in Fourier space.
This elementary calculation is the finite-rank shadow of the nested TBA
kernel manipulations used in principal chiral and Gross--Neveu models: the
Cartan matrix controls which auxiliary species couple, while the analytic
input lies in the physical scalar factors and in the thermodynamic string
hypothesis.

\section{Baxter TQ Relation}

For the \(XXX_{1/2}\) transfer eigenvalue
\eqref{eq:aba-xxx-transfer-eigenvalue}, define the Baxter polynomial
\[
  Q(u)=\prod_{j=1}^M(u-u_j).
\]
Then the eigenvalue equation can be rewritten as the finite-difference
relation
\begin{equation}
  \Lambda(u)Q(u)
  =
  \left(u+\frac{\ii}{2}\right)^L Q(u-\ii)
  +
  \left(u-\frac{\ii}{2}\right)^L Q(u+\ii).
  \label{eq:xxx-baxter-tq}
\end{equation}

Let \(Q(u)=\prod_j(u-u_j)\) have simple roots.  Equation
\eqref{eq:xxx-baxter-tq} is regular at the zeros of \(Q\) precisely when
the roots \(u_j\) obey the \(XXX_{1/2}\) Bethe equations
\eqref{eq:aba-xxx-bethe-equations}.  Evaluating the numerator at
\(u=u_j\) gives
\[
  \left(u_j+\frac{\ii}{2}\right)^L Q(u_j-\ii)
  +
  \left(u_j-\frac{\ii}{2}\right)^L Q(u_j+\ii)=0.
\]
Since
\[
  \frac{Q(u_j+\ii)}{Q(u_j-\ii)}
  =
  -\prod_{\substack{k=1\\ k\neq j}}^M
  \frac{u_j-u_k+\ii}{u_j-u_k-\ii},
\]
the displayed equality is precisely
\eqref{eq:aba-xxx-bethe-equations}.  Conversely, if the Bethe equations hold,
the numerator on the right side of \eqref{eq:xxx-baxter-tq} vanishes at every
root of \(Q\), so division by \(Q\) produces a polynomial \(\Lambda\).

\section{Baxter Casoratian and the Second Solution}
\label{sec:baxter-casoratian-second-solution}

The scalar \(TQ\) relation is a second-order finite-difference equation.  Its
two independent solutions carry the same information as a rank-two
Wronskian system.  This elementary point becomes structurally important in
higher rank and in the quantum spectral curve: a \(Q\)-system is not just a
list of Bethe polynomials, but a family of finite-difference solutions tied
together by Casoratian identities and boundary conditions.

Write the rank-one Baxter equation in the form
\begin{equation}
  \Lambda(u)Q(u)
  =
  a(u)Q(u-\ii)+d(u)Q(u+\ii),
  \label{eq:rank-one-baxter-general-ad}
\end{equation}
where \(a,d,\Lambda\) are meromorphic functions on a common shift domain.
For two solutions \(Q_1,Q_2\), define the half-shift Casoratian
\begin{equation}
  \mathcal W[Q_1,Q_2](u)
  =
  Q_1\!\left(u+\frac{\ii}{2}\right)
  Q_2\!\left(u-\frac{\ii}{2}\right)
  -
  Q_1\!\left(u-\frac{\ii}{2}\right)
  Q_2\!\left(u+\frac{\ii}{2}\right).
  \label{eq:baxter-casoratian-definition}
\end{equation}

\begin{lemma}[Baxter Casoratian transport]
\label{lem:baxter-casoratian-transport}
If \(Q_1\) and \(Q_2\) solve \eqref{eq:rank-one-baxter-general-ad} with the
same \(a,d,\Lambda\), then
\begin{equation}
  d(u)\,
  \mathcal W[Q_1,Q_2]\!\left(u+\frac{\ii}{2}\right)
  =
  a(u)\,
  \mathcal W[Q_1,Q_2]\!\left(u-\frac{\ii}{2}\right).
  \label{eq:baxter-casoratian-transport}
\end{equation}
At a point where \(a(u)d(u)\ne0\), vanishing of the Casoratian on one
adjacent half-shift is therefore transported to the other.
\end{lemma}

\begin{proof}
Multiply \eqref{eq:rank-one-baxter-general-ad} for \(Q_1\) by \(Q_2(u)\),
multiply the same equation for \(Q_2\) by \(Q_1(u)\), and subtract.  The
terms containing \(\Lambda(u)Q_1(u)Q_2(u)\) cancel, leaving
\[
  a(u)
  \left(Q_1(u-\ii)Q_2(u)-Q_2(u-\ii)Q_1(u)\right)
  +
  d(u)
  \left(Q_1(u+\ii)Q_2(u)-Q_2(u+\ii)Q_1(u)\right)
  =
  0 .
\]
The second bracket is
\[
  \mathcal W[Q_1,Q_2]\!\left(u+\frac{\ii}{2}\right),
\]
whereas the first bracket is
\[
  -\mathcal W[Q_1,Q_2]\!\left(u-\frac{\ii}{2}\right).
\]
Substitution gives \eqref{eq:baxter-casoratian-transport}.  If
\(a(u)d(u)\ne0\), the displayed equation is a nonzero scalar multiple
relation between the two adjacent Casoratians, so either one vanishes exactly
when the other does.
\end{proof}

For the homogeneous \(XXX_{1/2}\) normalization of
\eqref{eq:xxx-baxter-tq},
\[
  a(u)=\left(u+\frac{\ii}{2}\right)^L,\qquad
  d(u)=\left(u-\frac{\ii}{2}\right)^L,
\]
so the Casoratian obeys
\begin{equation}
  \frac{\mathcal W(u+\ii/2)}{\mathcal W(u-\ii/2)}
  =
  \left(\frac{u+\ii/2}{u-\ii/2}\right)^L .
  \label{eq:xxx-casoratian-ratio}
\end{equation}
The polynomial Bethe solution is only one member of the two-dimensional
local solution space of the difference equation.  The companion solution is
usually meromorphic rather than polynomial; its poles and asymptotics encode
the boundary condition that selected the Bethe state.  Higher-rank
Wronskian \(Q_A\)'s generalize this rank-one Casoratian: the determinant
formula in the next section is the multi-solution version of
\eqref{eq:baxter-casoratian-definition}, and the QQ-relations are the
corresponding Pluecker identities.

\section{QQ-System and Bäcklund Flow}

For higher-rank systems it is more efficient to work with a family of
Q-functions indexed by subsets.  Let \(I=\{1,\ldots,N\}\).  For every subset
\(A\subset I\) introduce a meromorphic function \(Q_A(u)\), with
\(Q_\varnothing(u)=1\).  Write
\[
  f^{\pm}(u)=f\!\left(u\pm\frac{\ii}{2}\right).
\]
The rational \(\mathfrak{gl}_N\) QQ-system is
\begin{equation}
  Q_{Aab}(u)Q_A(u)
  =
  Q_{Aa}^{+}(u)Q_{Ab}^{-}(u)
  -
  Q_{Aa}^{-}(u)Q_{Ab}^{+}(u),
  \label{eq:qq-system}
\end{equation}
for \(a,b\notin A\), \(a\ne b\).  Here \(Aa=A\cup\{a\}\) and
\(Aab=A\cup\{a,b\}\).  Repeated use of \eqref{eq:qq-system} reconstructs
all multi-index Q-functions from a flag of one-index functions together with
normalization data.  Moving from \(Q_A\) to \(Q_{Aa}\) is the functional form
of a Bäcklund step: one removes one simple color direction from the auxiliary
problem, exactly as the nested monodromy removes one residual color space at
each level.

Let \(q_1(u),\ldots,q_N(u)\) be functions for which the determinants below
are defined.  For an ordered subset \(A=\{a_1<\cdots<a_m\}\), set
\[
  Q_A(u)
  =
  \det_{1\le r,s\le m}
  q_{a_r}^{[m+1-2s]}(u),
  \qquad
  f^{[k]}(u)=f\!\left(u+\frac{\ii k}{2}\right),
\]
and \(Q_\varnothing=1\).  Then the functions \(Q_A\) obey
\eqref{eq:qq-system}.
Fix \(A\) of size \(m\) and two indices \(a,b\notin A\).  Form the
\((m+2)\times(m+2)\) matrix whose columns are the shifted columns
corresponding to \(A,a,b\).  Apply the Desnanot--Jacobi identity to the
minor obtained by deleting the two columns associated with the outer shifts.
With the shifts ordered as \(m+2,m,\ldots,-m,-m-2\), the four minors are
precisely \(Q_{Aab}Q_A\), \(Q_{Aa}^{+}Q_{Ab}^{-}\), and
\(Q_{Aa}^{-}Q_{Ab}^{+}\), with the relative sign fixed by the column order.
This gives \eqref{eq:qq-system}.  The identity is purely determinant
algebra; no analytic assumption on the \(q_i\)'s is used beyond existence of
the displayed values.

\section{Bäcklund Elimination as Q-System Restriction}
\label{sec:backlund-elimination-qsystem-restriction}

The Bäcklund language used in nested Bethe ansatz should be understood as a
recursive statement about Q-system data, not as a separate dynamical
principle.  Let \(I=\{1,\ldots,N\}\) and let \(\{Q_A\}_{A\subset I}\) obey
\eqref{eq:qq-system}.  Choose a color \(c\in I\), and define an unnormalized
restricted family on \(I\setminus\{c\}\) by
\begin{equation}
  Q^{(c)}_A(u)=Q_{A\cup\{c\}}(u),
  \qquad A\subset I\setminus\{c\}.
  \label{eq:backlund-restricted-qsystem}
\end{equation}
Then \(Q^{(c)}\) obeys the same local QQ-relation on the smaller index set:
\begin{equation}
  Q^{(c)}_{Aab}(u)Q^{(c)}_A(u)
  =
  \bigl(Q^{(c)}_{Aa}\bigr)^{+}(u)
  \bigl(Q^{(c)}_{Ab}\bigr)^{-}(u)
  -
  \bigl(Q^{(c)}_{Aa}\bigr)^{-}(u)
  \bigl(Q^{(c)}_{Ab}\bigr)^{+}(u),
  \label{eq:backlund-restricted-qq-system}
\end{equation}
for \(a,b\notin A\cup\{c\}\), \(a\ne b\).  This is simply
\eqref{eq:qq-system} applied to the subset \(A\cup\{c\}\) and the two
remaining indices \(a,b\).  The point of recording it is not the algebraic
verification; the point is the interpretation.  A Bäcklund step removes one
auxiliary color direction by passing from the \(N\)-color Q-system to the
subsystem of minors that contain the eliminated color.  Repeating the
operation along a flag
\[
  I=I_N\supset I_{N-1}\supset\cdots\supset I_1\supset I_0=\varnothing,
  \qquad |I_r|=r,
\]
produces the chain of rank-reduced Q-systems that mirrors the recursive
nested monodromy of
Chapter~\ref{ch:integrable-nested-bethe-ansatz-matrix-bethe-yang}.

The adjective ``unnormalized'' in \eqref{eq:backlund-restricted-qsystem}
matters.  The restricted empty function is \(Q^{(c)}_\varnothing=Q_c\), not
the constant \(1\).  In a finite spin chain one often chooses a Q-gauge in
which the new empty function is normalized away on a domain where \(Q_c\) has
no zeros.  In an analytic T-system or quantum spectral-curve problem this
normalization is part of the analytic datum: zeros, cuts, asymptotics, and
large-\(u\) powers cannot be divided out without recording what has happened
to them.  Thus Bäcklund elimination is a local algebraic operation on
Q-systems, while the normalized physical presentation of the reduced problem
requires the gauge and analytic data of
Definition~\ref{def:analytic-t-system-datum}.

For \(N=3\), the first elimination already displays the nesting pattern.  If
\(c=3\), the restricted rank-two data are
\[
  Q^{(3)}_\varnothing=Q_3,\qquad
  Q^{(3)}_1=Q_{13},\qquad
  Q^{(3)}_2=Q_{23},\qquad
  Q^{(3)}_{12}=Q_{123}.
\]
The smaller QQ-relation is
\[
  Q_{123}Q_3=Q_{13}^{+}Q_{23}^{-}-Q_{13}^{-}Q_{23}^{+},
\]
which is the original QQ-relation with base subset \(\{3\}\).  In nested
Bethe ansatz this is the algebraic shadow of passing from the original
three-color auxiliary problem to the residual two-color one; the Bethe roots
of the reduced problem are zeros of the Q-functions that survive the
restriction, together with the gauge convention used to interpret them as
polynomials.

\section{Fusion and the Hirota T-System}

The transfer matrix with a fundamental auxiliary space is only the first
member of a fused family.  In a rational \(SU(N)\) chain, one obtains
commuting transfer matrices \(T_{a,s}(u)\) by taking auxiliary spaces whose
highest weights are rectangular Young diagrams: \(a\) rows and \(s\) columns,
with \(1\le a\le N-1\) and \(s\ge0\).  Fusion is implemented by projecting
products of fundamental monodromies at shifted spectral parameters onto the
corresponding irreducible auxiliary component.  The Yang--Baxter equation
ensures that the projected transfer matrices commute.

The local functional identity is the Hirota equation
\begin{equation}
  T_{a,s}^{+}(u)T_{a,s}^{-}(u)
  =
  T_{a+1,s}(u)T_{a-1,s}(u)
  +
  T_{a,s+1}(u)T_{a,s-1}(u),
  \label{eq:nested-hirota-t-system}
\end{equation}
with boundary data \(T_{0,s}=T_{a,0}=1\) in the simplest normalization and
\(T_{N,s}\) fixed by the quantum determinant.  The precise boundary factors
depend on twist, inhomogeneity, and representation; the local bilinear
identity is the invariant content.

Suppose the fused eigenvalues \(T_{a,s}\) are written in a Wronskian gauge as
minors of a matrix of shifted Q-functions.  Then
\eqref{eq:nested-hirota-t-system} follows from the Pluecker relation among
adjacent minors.
In the Wronskian gauge, \(T_{a,s}\) is an \(a\times a\) minor whose columns
are shifts separated by \(\ii\).  The shifts of \(T_{a,s}^{+}\) and
\(T_{a,s}^{-}\) differ by moving the left and right boundary columns of this
minor outward.  The Desnanot--Jacobi identity for the corresponding
\((a+1)\times(a+1)\) matrix states that the product of the two adjacent
\(a\times a\) minors equals the product of the inner and outer minors plus
the product of the two side minors.  These four minors are, respectively,
\(T_{a,s}^{+}T_{a,s}^{-}\),
\(T_{a+1,s}T_{a-1,s}\), and \(T_{a,s+1}T_{a,s-1}\).  This is exactly
\eqref{eq:nested-hirota-t-system}.

The associated Y-functions are gauge-invariant ratios of neighboring
T-functions:
\begin{equation}
  Y_{a,s}(u)
  =
  \frac{T_{a,s+1}(u)T_{a,s-1}(u)}
       {T_{a+1,s}(u)T_{a-1,s}(u)}.
  \label{eq:t-to-y-definition}
\end{equation}
Using Hirota gives the local Y-system identity
\begin{equation}
  Y_{a,s}^{+}(u)Y_{a,s}^{-}(u)
  =
  \frac{(1+Y_{a,s+1})(1+Y_{a,s-1})}
       {(1+Y_{a+1,s}^{-1})(1+Y_{a-1,s}^{-1})},
  \label{eq:hirota-to-y-system}
\end{equation}
where all functions are evaluated at \(u\) unless shifted.  This is an
algebraic identity once the same Hirota gauge is used at the four adjacent
nodes; analytic continuation, strip domains, source terms, and asymptotics
are additional model-specific data.

\section{T-Gauge and Analytic T-System Specifications}

The Hirota equation is not by itself a spectral problem.  It is a local
bilinear relation on a two-dimensional lattice of functions.  A transfer
matrix, a relativistic TBA system, or a quantum spectral curve is obtained
only after one adds the boundary conditions and analytic constraints that
select a particular solution.  The first such constraint is the gauge freedom
of the T-system.

\begin{definition}[T-gauge transformation]
\label{def:t-system-gauge-transformation}
Let \(g_1,g_2,g_3,g_4\) be nowhere-vanishing meromorphic functions on a shift
domain containing the arguments used below.  For a family \(T_{a,s}(u)\), set
\begin{equation}
  \mathcal G_{a,s}(u)
  =
  g_1^{[a+s]}(u)\,
  g_2^{[a-s]}(u)\,
  g_3^{[-a+s]}(u)\,
  g_4^{[-a-s]}(u),
  \qquad
  \widetilde T_{a,s}(u)=\mathcal G_{a,s}(u)T_{a,s}(u).
  \label{eq:t-system-gauge-transformation}
\end{equation}
This transformation is called a T-gauge transformation on the chosen
\((a,s)\)-domain.
\end{definition}

The four factors in \(\mathcal G_{a,s}\) are arranged so that every monomial
in the Hirota equation receives the same multiplier:
\begin{equation}
  \mathcal G_{a,s}^{+}\mathcal G_{a,s}^{-}
  =
  \mathcal G_{a+1,s}\mathcal G_{a-1,s}
  =
  \mathcal G_{a,s+1}\mathcal G_{a,s-1}.
  \label{eq:t-gauge-hirota-cocycle}
\end{equation}
For example, the \(g_1\) arguments in the first product are
\([a+s+1]\) and \([a+s-1]\); the \(g_1\) arguments in the second and third
products are the same two shifts in the opposite lattice directions.  The
same check applies to \(g_2,g_3,g_4\).  Thus a T-gauge transformation maps a
solution of \eqref{eq:nested-hirota-t-system} to another solution.  It also
leaves the Y-functions \eqref{eq:t-to-y-definition} invariant, because the
four \(g_i\)-shift pairs in the numerator and denominator cancel separately.
Consequently the Y-system is a gauge-invariant shadow of the T-system, while
the transfer matrices themselves retain gauge-dependent boundary factors.

This distinction matters in applications.  The local Hirota relation says
that adjacent rectangular representations, or adjacent Wronskian minors,
fit together consistently.  It does not specify:
\begin{enumerate}
\item the allowed \((a,s)\)-domain;
\item the boundary factors \(T_{0,s}\), \(T_{N,s}\), \(T_{a,0}\), and their
      zeros and poles;
\item the strip of meromorphy or cut structure in the spectral parameter;
\item asymptotic behavior at infinity or at distinguished singular points;
\item reality, crossing, twist, and quantum-determinant constraints;
\item source terms introduced by physical particles, holes, defects, or
      wrapping corrections.
\end{enumerate}
These items are not decorative.  They are the analytic and boundary conditions
that turn the Hirota
identity into a spectral problem with a discrete set, or a controlled
continuum family, of solutions.  Two T-systems related by
\eqref{eq:t-system-gauge-transformation} describe the same Y-system, but they
may place boundary divisors in different T-functions.  A statement about
polynomiality, zero locations, or large-\(u\) normalization is therefore
meaningful only after the gauge has been fixed.

\begin{definition}[Analytic T-system specification]
\label{def:analytic-t-system-datum}
An analytic T-system specification on a lattice domain
\(\mathfrak D\subset\mathbb Z^2\)
consists of:
\begin{enumerate}
\item meromorphic functions \(T_{a,s}\) on a specified shift-stable spectral
      domain, for \((a,s)\in\mathfrak D\), satisfying Hirota at interior
      lattice points;
\item a T-gauge equivalence class, together with a chosen gauge when
      polynomiality or asymptotics are asserted;
\item boundary functions and quantum-determinant factors on
      \(\partial\mathfrak D\);
\item the allowed zero, pole, branch-cut, and asymptotic conditions;
\item the map from this analytic specification to physical charges or transfer
      eigenvalues.
\end{enumerate}
\end{definition}

In finite spin chains the analytic specification is often polynomial: the
inhomogeneities, twist, representation labels, and quantum determinant fix
the boundary factors, and Bethe equations enforce cancellation of unwanted
poles.  In relativistic integrable QFT, the same local Hirota/Y-system
algebra must be supplemented by strip analyticity, source terms, and
large-rapidity asymptotics before it becomes a TBA problem.  In planar
gauge-theory integrability, the quantum spectral curve replaces the finite
polynomial condition by a Riemann--Hilbert type problem: sheets, cuts,
asymptotics, gluing conditions, and quantization of charges.  The determinant
identities in this chapter supply the local algebraic skeleton; the analytic
T-system specification turns that skeleton into a physical spectral problem.

\section{Q-Operators}

A \(Q\)-operator is an operator-valued function \(Q(u)\) commuting with the
transfer matrix and satisfying an operator \(TQ\) relation whose eigenvalue
version is \eqref{eq:xxx-baxter-tq}.  Operator construction requires an
auxiliary representation in which the trace, or a regulated trace, is
defined.  In rank one this can be described as follows.

\begin{construction}[Rank-one Baxter operator construction]
\label{con:rank-one-q-operator-datum}
Let \(\mathcal A\) be an auxiliary representation of the RTT algebra for the
same \(R\)-matrix, not necessarily finite dimensional.  Let
\(\mathcal L_{\mathcal A,j}(u)\in
\operatorname{End}(\mathcal A\otimes\C^2_j)\) satisfy the mixed RTT relation
with the fundamental \(L\)-operator.  If the regularized auxiliary trace
\[
  Q(u)=\operatorname{Tr}_{\mathcal A}^{\rm reg}
  \bigl(
    \mathcal L_{\mathcal A,L}(u)\cdots
    \mathcal L_{\mathcal A,1}(u)
  \bigr)
\]
exists and is cyclic on the products used in the RTT proof, then
\([Q(u),t(v)]=0\).  If the tensor product of the fundamental auxiliary space
with \(\mathcal A\) has a short exact sequence whose quotient and submodule
shift the spectral parameter by \(\pm\ii\), the transfer-matrix trace over
that sequence gives the operator \(TQ\) relation.
\end{construction}

\paragraph{Exact-sequence derivation of an operator \(TQ\) relation.}
Assume the data of Construction~\ref{con:rank-one-q-operator-datum}.  Assume
further that, for every spectral parameter \(u\), there is a short exact
sequence of auxiliary RTT modules
\begin{equation}
  0\longrightarrow
  \mathcal A(u-\ii)
  \stackrel{\iota}{\longrightarrow}
  \C^2(u)\otimes\mathcal A(u)
  \stackrel{\pi}{\longrightarrow}
  \mathcal A(u+\ii)
  \longrightarrow0 ,
  \label{eq:q-operator-short-exact-sequence}
\end{equation}
and that the corresponding monodromy on
\(\C^2(u)\otimes\mathcal A(u)\otimes\mathcal H_L\) is block triangular in a
basis adapted to \(\iota,\pi\), with diagonal blocks
\(a(u)\,M_{\mathcal A}(u-\ii)\) and
\(d(u)\,M_{\mathcal A}(u+\ii)\).  Then
\begin{equation}
  t(u)\,Q(u)
  =
  a(u)Q(u-\ii)+d(u)Q(u+\ii)
  \label{eq:operator-tq-from-exact-sequence}
\end{equation}
as operators on the quantum chain.

The product \(t(u)Q(u)\) is the regularized trace over
\(\C^2(u)\otimes\mathcal A(u)\) of the product monodromy, because the mixed
RTT relation permits the fundamental and auxiliary monodromies to be ordered
inside one common auxiliary trace.  Adapt the auxiliary vector space to the
short exact sequence \eqref{eq:q-operator-short-exact-sequence}.  The
monodromy matrix is block triangular, hence its trace is the sum of the
traces of the two diagonal blocks.  The scalar diagonal factors are exactly
the vacuum propagation factors \(a(u)\) and \(d(u)\), while the remaining
block traces are the regularized traces defining \(Q(u-\ii)\) and
\(Q(u+\ii)\).  Cyclicity of the regularized trace is the only analytic input
in this finite-chain argument.  This gives
\eqref{eq:operator-tq-from-exact-sequence}.

The construction displays the logical ingredients rather than hiding them in
the phrase ``Baxter operator'': a mixed RTT relation, a trace functional with
cyclicity, and an exact sequence of auxiliary representations.  In rational
finite chains one may realize \(\mathcal A\) as a limiting highest-weight
module.  In trigonometric rank-one chains one uses a q-oscillator Fock
module.  For \(q\in\C^\times\) not a root of unity, one convenient
q-oscillator algebra is generated by \(D,\mathsf a,\mathsf a^\dagger\) with
\[
  q^D\mathsf a q^{-D}=q^{-1}\mathsf a,\qquad
  q^D\mathsf a^\dagger q^{-D}=q\mathsf a^\dagger,\qquad
  \mathsf a^\dagger\mathsf a=1-q^{2D},\qquad
  \mathsf a\mathsf a^\dagger=1-q^{2D+2}.
\]
On the Fock module \(\mathcal F=\bigoplus_{n\ge0}\C\ket n\),
\[
  D\ket n=n\ket n,\qquad
  \mathsf a\ket n=(1-q^{2n})\ket{n-1},\qquad
  \mathsf a^\dagger\ket n=\ket{n+1}.
\]
This convention is related to the Borel \(q\)-oscillator generators
\(\mathcal H,\mathcal E_\pm\) by
\[
  \mathcal H=2D,\qquad
  \mathcal E_+=\frac{\mathsf a^\dagger}{q-q^{-1}},\qquad
  \mathcal E_-=\frac{\mathsf a}{q-q^{-1}}.
\]
Indeed
\begin{equation}
  [\mathcal H,\mathcal E_\pm]=\pm 2\mathcal E_\pm,\qquad
  q\mathcal E_+\mathcal E_-
  -q^{-1}\mathcal E_-\mathcal E_+
  =\frac{1}{q-q^{-1}},
  \label{eq:qosc-borel-generators}
\end{equation}
where the second identity follows from
\[
  q\mathsf a^\dagger\mathsf a-q^{-1}\mathsf a\mathsf a^\dagger
  =
  q(1-q^{2D})-q^{-1}(1-q^{2D+2})
  =q-q^{-1}.
\]

\paragraph{The local trigonometric \(q\)-oscillator \(L\)-operator.}
Let \(V=\C e_+\oplus\C e_-\).  In the six-vertex normalization used in
this paragraph, \(R(x)\in\operatorname{End}(V\otimes V)\) has non-zero entries
\begin{align}
  R^{++}_{++}(x)&=R^{--}_{--}(x)=q+q^{-1}x,&
  R^{+-}_{+-}(x)&=R^{-+}_{-+}(x)=1+x,\nonumber\\
  R^{+-}_{-+}(x)&=q-q^{-1},&
  R^{-+}_{+-}(x)&=-(q-q^{-1})x .
  \label{eq:qosc-six-vertex-r-normalization}
\end{align}
The q-oscillator auxiliary \(L\)-operator is the following element of
\(\operatorname{End}(V\otimes\mathcal F)\), written as a \(2\times2\) matrix
in the ordered basis \(e_+,e_-\):
\begin{equation}
  L_{\mathcal F}^{(+)}(x)=
  \begin{pmatrix}
    q^D&\mathsf a\\
    -x\,\mathsf a^\dagger&q^{-D}+xq^{D+1}
  \end{pmatrix}.
  \label{eq:qosc-l-operator-plus}
\end{equation}
Equivalently, before choosing the Fock realization,
\begin{align}
  L_{\mathcal F}^{(+)}(x)
  &=
  q^{\mathcal H\otimes\sigma^z/2}
  +(q-q^{-1})(\mathcal E_-\otimes\sigma^+
      -x\mathcal E_+\otimes\sigma^-)
  \nonumber\\
  &\quad
  +x(q-q^{-1})q^{-\mathcal H/2}
  (\mathcal E_+\mathcal E_--\mathcal E_-\mathcal E_+)
  \otimes\sigma^-\sigma^+ ,
  \label{eq:qosc-l-operator-borel-form}
\end{align}
with \(\sigma^z e_\pm=\pm e_\pm\), \(\sigma^+e_-=e_+\),
\(\sigma^-e_+=e_-\).  Substituting
\eqref{eq:qosc-borel-generators} gives
\eqref{eq:qosc-l-operator-plus}; the bottom-right entry uses
\[
  \mathcal E_+\mathcal E_- - \mathcal E_-\mathcal E_+
  =
  \frac{q^{2D+1}}{q-q^{-1}}.
\]

\begin{proposition}[Local \(q\)-oscillator RLL identity]
\label{prop:qosc-local-rll-identity}
For \(q\in\C^\times\) not a root of unity, the operators
\eqref{eq:qosc-six-vertex-r-normalization} and
\eqref{eq:qosc-l-operator-plus} satisfy
\begin{equation}
  R_{12}(-x/y)\,
  L^{(+)}_{\mathcal F,13}(x)\,
  L^{(+)}_{\mathcal F,23}(y)
  =
  L^{(+)}_{\mathcal F,23}(y)\,
  L^{(+)}_{\mathcal F,13}(x)\,
  R_{12}(-x/y)
  \label{eq:qosc-local-rll}
\end{equation}
as an identity in
\(\operatorname{End}(V\otimes V\otimes\mathcal F)\).
\end{proposition}

\begin{proof}
Write \(N=q^D\).  The q-oscillator relations are
\begin{equation}
  N\mathsf a=q^{-1}\mathsf aN,\qquad
  N\mathsf a^\dagger=q\mathsf a^\dagger N,\qquad
  \mathsf a^\dagger\mathsf a=1-N^2,\qquad
  \mathsf a\mathsf a^\dagger=1-q^2N^2 .
  \label{eq:qosc-local-rll-relations}
\end{equation}
Both sides of \eqref{eq:qosc-local-rll} are finite sums of monomials in
\(N^{\pm1},\mathsf a,\mathsf a^\dagger\).  Moving every \(N^{\pm1}\)
to the middle and reducing every adjacent
\(\mathsf a^\dagger\mathsf a\) and
\(\mathsf a\mathsf a^\dagger\) by
\eqref{eq:qosc-local-rll-relations} gives a normal form with at most one
creation or annihilation operator on either side of a Laurent polynomial in
\(N\).  Therefore it is enough to compare the four spin-input sectors
\(e_{\epsilon_1}\otimes e_{\epsilon_2}\otimes\ket n\), \(n\ge0\).

Let
\begin{align*}
  N_n&=q^n,&
  \alpha_n&=1-q^{2n},&
  A_n&=N_n^2-1,&
  \Delta&=x-q^2y,\\
  B_x&=N_n^2qx+1,&
  B_y&=N_n^2qy+1,&
  G_y&=N_n^2y+q .
\end{align*}
For the mixed sector below also write
\begin{align*}
  C_{+-}&=N_n^2q^2y^2-N_n^2xy-q^2xy-qx+qy+xy,\\
  D_{+-}&=N_n^2q^3y-N_n^2qx+q^2+qx-qy-1 .
\end{align*}
In the sector \(e_+\otimes e_+\), both sides of
\eqref{eq:qosc-local-rll} act as
\begin{align}
  e_+\otimes e_+\otimes\ket n
  \mapsto\;&
  -\frac{\Delta}{qy}N_n^2\,e_+\otimes e_+\otimes\ket n
  +\frac{\Delta}{q}N_n\,e_+\otimes e_-\otimes\ket{n+1}
  \nonumber\\
  &+\frac{x\Delta}{y}N_n\,e_-\otimes e_+\otimes\ket{n+1}
  -\frac{x\Delta}{q}\,e_-\otimes e_-\otimes\ket{n+2}.
  \label{eq:qosc-rll-plus-plus-sector}
\end{align}
In the sector \(e_-\otimes e_-\), applying the same reductions gives
\begin{align}
  e_-\otimes e_-\otimes\ket n
  \mapsto\;&
  \frac{(q-N_n)(N_n+q)A_n\Delta}{q^3y}\,
    e_+\otimes e_+\otimes\ket{n-2}
  +\frac{A_nG_y\Delta}{qN_ny}\,
    e_+\otimes e_-\otimes\ket{n-1}
  \nonumber\\
  &+\frac{A_n\Delta B_x}{qN_ny}\,
    e_-\otimes e_+\otimes\ket{n-1}
  -\frac{\Delta B_xB_y}{qN_n^2y}\,
    e_-\otimes e_-\otimes\ket n .
  \label{eq:qosc-rll-minus-minus-sector}
\end{align}
The terms with \(\ket{n-1}\) or \(\ket{n-2}\) vanish automatically when
the displayed ket is outside the Fock module, because
\(\alpha_0=0\).  The mixed sectors are the only places where the
off-diagonal six-vertex weights enter.  For \(e_+\otimes e_-\), both sides
give
\begin{align}
  &-\frac{\Delta\alpha_nN_n}{q^2y}\,
    e_+\otimes e_+\otimes\ket{n-1}
  +\frac{C_{+-}}{qy}\,
    e_+\otimes e_-\otimes\ket n
  \nonumber\\
  &\quad
  +\frac{xD_{+-}}{qy}\,
    e_-\otimes e_+\otimes\ket n
  +\frac{x\Delta B_y}{qN_ny}\,
    e_-\otimes e_-\otimes\ket{n+1}.
  \label{eq:qosc-rll-plus-minus-sector}
\end{align}
The \(e_-\otimes e_+\) formula is obtained from
\eqref{eq:qosc-rll-plus-minus-sector} by interchanging
\((x,e_1)\leftrightarrow(y,e_2)\) together with the asymmetric
normalization of \(R^{-+}_{+-}\) in
\eqref{eq:qosc-six-vertex-r-normalization}; the resulting coefficients
are exactly those obtained by expanding the two sides.  These four sector
computations exhaust the basis of \(V\otimes V\otimes\mathcal F\), hence
\eqref{eq:qosc-local-rll} follows.
\end{proof}

The regularized trace is first defined by
\[
  \operatorname{Tr}^{\rm reg}_{\mathcal F}(X)
  =
  \operatorname*{mer.cont.}_{z=1}\operatorname{Tr}_{\mathcal F}(z^D X),
  \qquad |z|<1,
\]
whenever the meromorphic continuation exists and is finite after the scalar
normalization of the \(Q\)-operator has been fixed.  Proposition
\ref{prop:qosc-local-rll-identity} supplies the local algebraic input for
the trigonometric rank-one \(Q\)-operator.  The remaining analytic inputs
are separate: cyclicity of the regularized trace on the monodromy products
and the short exact sequence \eqref{eq:q-operator-short-exact-sequence}.
Once these are checked, the operator identity
\eqref{eq:operator-tq-from-exact-sequence} is a consequence of
block-triangular algebra plus the stated trace regularization.  The
eigenvalue statement used in this chapter is the consequence of this
construction, not its replacement.

\section{Separation of Variables}

Separation of variables uses a different complete set of commuting objects.
For the \(2\times2\) monodromy
\[
  T(u)=
  \begin{pmatrix}
    A(u)&B(u)\\
    C(u)&D(u)
  \end{pmatrix},
\]
RTT implies \([B(u),B(v)]=0\).  Suppose we are in a finite-dimensional
representation for which the commuting coefficients of \(B(u)\) are
simultaneously diagonalizable with simple joint spectrum.  On a joint
right or left eigenline the \(B\)-eigenvalue has the polynomial form
\[
  B(u)=B_0\prod_{\alpha=1}^{L-1}(u-x_\alpha).
\]
The numbers \(x_\alpha\) are separated coordinates.

\paragraph{Local RTT shift algebra.}
The separated-coordinate shift is an operator statement before it is a
wavefunction equation.  In the rational RTT normalization compatible with
\eqref{eq:xxx-baxter-tq}, the components involving \(A,D\), and \(B\) are
understood as identities after multiplication by \(u-v\):
\begin{align}
  A(u)B(v)
  &=
  \frac{u-v-\ii}{u-v}\,B(v)A(u)
  +
  \frac{\ii}{u-v}\,B(u)A(v),
  \label{eq:sov-rtt-a-b-component}\\
  D(u)B(v)
  &=
  \frac{u-v+\ii}{u-v}\,B(v)D(u)
  -
  \frac{\ii}{u-v}\,B(u)D(v).
  \label{eq:sov-rtt-d-b-component}
\end{align}
Let \(\bra{\boldsymbol x}\) be a simultaneous left eigenvector of the
commuting coefficients of \(B\):
\[
  \bra{\boldsymbol x}B(v)
  =
  B_0\prod_{\beta=1}^{L-1}(v-x_\beta)\,\bra{\boldsymbol x}.
\]
Set \(u=x_\alpha\) in
\eqref{eq:sov-rtt-a-b-component}--\eqref{eq:sov-rtt-d-b-component}.  The
terms proportional to \(B(x_\alpha)\) vanish on the left eigenvector, so
\begin{align}
  \bra{\boldsymbol x}A(x_\alpha)B(v)
  &=
  B_0\,(v-x_\alpha+\ii)
  \prod_{\beta\ne\alpha}(v-x_\beta)\,
  \bra{\boldsymbol x}A(x_\alpha),
  \label{eq:sov-a-shifts-one-zero}\\
  \bra{\boldsymbol x}D(x_\alpha)B(v)
  &=
  B_0\,(v-x_\alpha-\ii)
  \prod_{\beta\ne\alpha}(v-x_\beta)\,
  \bra{\boldsymbol x}D(x_\alpha).
  \label{eq:sov-d-shifts-one-zero}
\end{align}
Thus, whenever the resulting covector is nonzero and the joint spectrum is
simple, \(\bra{\boldsymbol x}A(x_\alpha)\) lies on the eigenline with the
single coordinate \(x_\alpha\) replaced by \(x_\alpha-\ii\), while
\(\bra{\boldsymbol x}D(x_\alpha)\) lies on the eigenline with
\(x_\alpha\) replaced by \(x_\alpha+\ii\).  A choice of SoV covector
normalization fixes the scalar coefficients.  In the normalization used in
the scalar Baxter equation,
\begin{equation}
  \bra{\boldsymbol x}A(x_\alpha)
  =
  a(x_\alpha)\bra{\boldsymbol x-\ii e_\alpha},
  \qquad
  \bra{\boldsymbol x}D(x_\alpha)
  =
  d(x_\alpha)\bra{\boldsymbol x+\ii e_\alpha}.
  \label{eq:sov-covector-shift-normalization}
\end{equation}
Equations
\eqref{eq:sov-a-shifts-one-zero}--\eqref{eq:sov-covector-shift-normalization}
are the finite-chain construction behind the separated Baxter equation
below.  The shift is not an analogy with classical separated coordinates; it
is the residue of the RTT algebra at a zero of \(B\).

\paragraph{Finite-chain SoV reduction under simple-spectrum hypotheses.}
Assume the coefficients of \(B(u)\) have a simple joint eigenbasis
\(\{\bra{\boldsymbol x}\}\) normalized as in
\eqref{eq:sov-covector-shift-normalization}.  Then a transfer-matrix
eigenvector with separated wavefunction
\(\Psi(\boldsymbol x)=\langle\boldsymbol x|\Psi\rangle\) satisfies, for each
\(\alpha=1,\ldots,L-1\), a one-variable Baxter equation
\begin{equation}
  \Lambda(x_\alpha)\Psi(\boldsymbol x)
  =
  a(x_\alpha)\Psi(\boldsymbol x-\ii e_\alpha)
  +
  d(x_\alpha)\Psi(\boldsymbol x+\ii e_\alpha),
  \label{eq:sov-coordinate-baxter}
\end{equation}
where \(a(u)\) and \(d(u)\) are the vacuum diagonal eigenvalue functions of
\(A(u)\) and \(D(u)\), and \(e_\alpha\) shifts only the coordinate
\(x_\alpha\).

Act with \(t(u)=A(u)+D(u)\) on a transfer eigenvector \(\Psi\) and evaluate
the left covector equation at \(u=x_\alpha\).  The two summands are evaluated
by \eqref{eq:sov-covector-shift-normalization}, giving exactly
\eqref{eq:sov-coordinate-baxter}.  The simple-spectrum hypothesis ensures
that the shifted covectors are identified by their separated coordinates
rather than by a nontrivial Jordan block or multiplicity space.

When both the algebraic Bethe ansatz and SoV are valid, the Bethe polynomial
\(Q(u)\) is the one-variable factor whose values solve
\eqref{eq:sov-coordinate-baxter} at every separated coordinate.  The
pseudovacuum construction builds eigenvectors from zeros of \(Q\); SoV builds
the same spectral problem by diagonalizing \(B(u)\).  Their equality is not a
slogan about roots: it is a comparison statement between two bases of the same
finite-dimensional representation.

\paragraph{Finite-chain ABA--\(Q\)--SoV comparison under explicit hypotheses.}
For a finite rational \(SU(2)\) chain with generic diagonal twist and generic
inhomogeneities, assume the following finite-dimensional properties:
\begin{enumerate}
\item the transfer matrix has simple spectrum;
\item the algebraic Bethe ansatz produces one nonzero Bethe vector for every
transfer eigenvalue;
\item a Baxter \(Q\)-operator satisfying the operator \(TQ\) relation exists;
\item the coefficients of \(B(u)\) have a simple SoV spectrum.
\end{enumerate}
Then the ABA Bethe roots, the zeros of the \(Q\)-operator eigenvalue, and the
one-variable SoV Baxter wavefunction determine the same transfer-matrix
eigenline.

The justification uses only these finite-chain inputs.  By the operator
\(TQ\) relation, the eigenvalue \(Q_\Psi(u)\) of \(Q(u)\) on a
transfer eigenvector \(\Psi\) satisfies \eqref{eq:xxx-baxter-tq} with the
same \(\Lambda(u)\).  Regularity of the scalar \(TQ\) relation at the simple
zeros of \(Q_\Psi\) gives the ABA Bethe equations.
Assumption 2 then gives an ABA vector with the same transfer eigenvalue
\(\Lambda\).  Simplicity of the transfer spectrum identifies the two vectors
up to scalar.  The SoV wavefunction satisfies
\eqref{eq:sov-coordinate-baxter}; the polynomial \(Q_\Psi\) gives a solution
of each one-coordinate equation by evaluation at \(x_\alpha\).  Simplicity of
the SoV joint spectrum identifies the resulting wavefunction with the same
eigenline.  Thus all three constructions label the same line, under the
stated hypotheses.

\section{The SoV Measure Is Part of the Construction}

The word ``measure'' in finite-chain SoV does not mean a Borel measure on a
classical configuration space.  It is the weight in the resolution of the
identity for a finite-dimensional joint spectral problem.  Let
\(\mathcal X\) be the finite joint spectrum of the commuting coefficients of
\(B(u)\), and let \(\ket{\boldsymbol x}\),
\(\bra{\widetilde{\boldsymbol x}}\) be right and left eigenvectors normalized
so that
\[
  \braket{\widetilde{\boldsymbol x}}{\boldsymbol y}=0
  \qquad(\boldsymbol x\ne\boldsymbol y).
\]
If the joint spectrum is simple and complete, define
\[
  \mu(\boldsymbol x)
  =
  \braket{\widetilde{\boldsymbol x}}{\boldsymbol x}^{-1}.
\]
Then
\begin{equation}
  \id_{\mathcal H}
  =
  \sum_{\boldsymbol x\in\mathcal X}
  \mu(\boldsymbol x)
  \ket{\boldsymbol x}\bra{\widetilde{\boldsymbol x}} .
  \label{eq:sov-resolution-identity}
\end{equation}
For two vectors \(\Phi,\Psi\), their scalar product is therefore
\[
  \braket{\Phi}{\Psi}
  =
  \sum_{\boldsymbol x\in\mathcal X}
  \mu(\boldsymbol x)\,
  \braket{\Phi}{\boldsymbol x}
  \braket{\widetilde{\boldsymbol x}}{\Psi},
\]
with the obvious conjugations when the left basis is chosen as the Hermitian
dual of the right basis in a unitary representation.  In concrete
inhomogeneous spin chains this weight can be written as a determinant or
Vandermonde-type product.  That explicit formula is representation-dependent,
but \eqref{eq:sov-resolution-identity} is the invariant finite-dimensional
statement.

Under the simple-spectrum SoV hypotheses above, suppose a
function \(Q(u)\) solves the one-variable Baxter equation at every separated
coordinate value appearing in \(\mathcal X\).  Then
\[
  \Psi_Q(\boldsymbol x)=\prod_{\alpha=1}^{L-1}Q(x_\alpha)
\]
solves the separated transfer-matrix eigenvalue equations on
\(\mathcal X\).  Conversely, if the separated spectrum is a Cartesian product
of one-coordinate spectra and the one-coordinate Baxter equation has a
one-dimensional solution space after the boundary conditions are imposed,
then every separated eigenfunction factorizes in this form.
Equation~\eqref{eq:sov-coordinate-baxter} shifts only one coordinate.  If
\(\Psi_Q=\prod_\beta Q(x_\beta)\), the common factors
\(\prod_{\beta\ne\alpha}Q(x_\beta)\) cancel from the \(\alpha\)-th separated
equation, leaving exactly the one-variable Baxter equation for \(Q\) at
\(x_\alpha\).  Conversely, on a Cartesian separated spectrum the equations
for different coordinates commute and act on distinct variables.  The stated
one-dimensionality of the one-coordinate solution space fixes each factor up
to a scalar; the product of those scalars is one overall normalization of
the eigenvector.

\section{Continuum Status}

The finite-chain comparison theorem does not close the continuum QFT problem.
For a relativistic integrable QFT one must additionally prove that
finite-volume energy levels are approximated by matrix Bethe--Yang equations
in the large-\(L\) regime, that the nested string hypothesis used in
Chapter~\ref{ch:integrable-nested-bethe-ansatz-matrix-bethe-yang} is valid
for the thermodynamic states under consideration, and that the integral
operators in the TBA equations have the required analytic and Fredholm
properties.  These are model-dependent mathematical statements.  They are not
part of the finite RTT algebra, and this monograph keeps them separate from
the algebraic construction.

\section{Interface with Planar Gauge-Theory Integrability}

In the planar \(\mathcal N=4\) SYM volume, magnons have a non-relativistic
dispersion on a spin chain, wrapping effects arise from trace cyclicity and
finite operator length, and the quantum spectral curve replaces the
finite-rank Baxter polynomial by a system of analytic \(Q\)-functions.  The
algebraic bridge is now explicit: the Q-system
\eqref{eq:qq-system}, the Hirota equation
\eqref{eq:nested-hirota-t-system}, and the \(TQ\)/SoV comparison theorems of
this chapter are the finite-rank prototypes.  Planar gauge theory changes the
analytic domain, asymptotics, grading, and boundary conditions; it does not
change the local determinant identities.
