\chapter{Thermodynamic Bethe Ansatz}
\label{ch:integrable-thermodynamic-bethe-ansatz}

\ConventionsLorentzian

The thermodynamic Bethe ansatz turns the exact two-body scattering data of a
massive integrable theory into finite-temperature or finite-size
thermodynamics.  The construction in this chapter is stated first for a
diagonal elastic \(S\)-matrix.  Matrix-valued scattering, bound-state strings,
and nested Bethe ansatz data require additional species and auxiliary-root
bookkeeping but obey the same variational logic once the allowed densities
are specified.

\begin{definition}[Diagonal massive TBA specification]
\label{def:diagonal-massive-tba-datum}
A diagonal massive thermodynamic Bethe ansatz specification consists of the following
objects.
\begin{enumerate}[label=\textup{(\roman*)}]
\item A finite species set \(\mathsf A\), with a mass
      \(m_a>0\) for each \(a\in\mathsf A\).
\item Diagonal two-body scattering amplitudes
      \(S_{ab}(\theta)\in U(1)\) for real rapidity difference
      \(\theta\), with a declared real-axis phase branch
      \[
        S_{ab}(\theta)=\exp\bigl(\ii\delta_{ab}(\theta)\bigr)
      \]
      on each rapidity interval used in the thermodynamic calculation.  The
      logarithmic derivative kernel is
      \[
        \varphi_{ab}(\theta)
        =
        \frac{1}{2\pi}\delta'_{ab}(\theta)
        =
        \frac{1}{2\pi\ii}\frac{\dd}{\dd\theta}\log S_{ab}(\theta)
      \]
      as a locally integrable function or distribution.  The branch convention
      is part of the specification whenever \(S_{ab}\) has constant signs, poles, or
      CDD factors crossing the chosen real-axis chart.
\item Circle quantization shifts
      \(\nu_a\in\mathbb R/\mathbb Z\), encoding the statistics and boundary
      convention for each species.
\item A thermodynamic root-density statement: for circle circumference
      \(L\to\infty\), the Bethe--Yang solutions whose rapidities stay in a
      fixed compact interval have empirical measures converging to particle
      densities \(\rho_a(\theta)\dd\theta\) and hole densities
      \(\rho_a^h(\theta)\dd\theta\), and the differentiated Bethe--Yang map
      gives the density constraint
      \[
        \rho_a+\rho_a^h
        =
        \frac{m_a}{2\pi}\cosh\theta
        +
        \sum_{b\in\mathsf A}\varphi_{ab}\ast\rho_b .
      \]
\item The exclusion-statistics entropy convention
      \[
        s[\rho,\rho^h]
        =
        \sum_{a\in\mathsf A}\int d\theta\,
        \left[
        \rho_a^t\log\rho_a^t
        -\rho_a\log\rho_a
        -\rho_a^h\log\rho_a^h
        \right],
        \qquad
        \rho_a^t=\rho_a+\rho_a^h .
      \]
      This convention is the one matched to one occupied or empty Bethe
      quantum number in each microscopic rapidity cell.  Other generalized
      exclusion-statistics systems require the corresponding entropy and
      density matrix in place of this item.
\item A thermal or mirror-channel normalization.  At inverse temperature
      \(R>0\), the energy density coordinate is
      \(\sum_a\int d\theta\,m_a\cosh\theta\,\rho_a(\theta)\).  When the same
      specification is used for a finite-size ground-state energy, the chapter assumes
      the mirror-channel identification and the normalization of
      \(E_0(R)\) stated in Section~\ref{sec:tba-finite-size-ground-state}.
\end{enumerate}
\end{definition}

\section{Bethe--Yang Quantization}

For the diagonal specification of
Definition~\ref{def:diagonal-massive-tba-datum}, place the theory on a
spatial circle of circumference \(L\).  For \(n\) particles of masses \(m_a\),
rapidities \(\theta_j\), and diagonal scattering phases
\[
  S_{ab}(\theta)=\exp\!\bigl(\ii\delta_{ab}(\theta)\bigr),
\]
periodicity of the wavefunction of particle \(j\) gives the Bethe--Yang
equation
\[
  m_{a_j}L\sinh\theta_j
  +\sum_{k\neq j}\delta_{a_ja_k}(\theta_j-\theta_k)
  =
  2\pi I_j,
  \qquad I_j\in\mathbb Z+\nu_j .
\]
The shift \(\nu_j\) depends on statistics and boundary convention.  The
one-particle energy and momentum are
\[
  E_a(\theta)=m_a\cosh\theta,\qquad
  p_a(\theta)=m_a\sinh\theta .
\]

\section{Density Equations}

In the thermodynamic limit \(L\to\infty\), introduce particle and hole
densities \(\rho_a(\theta)\) and \(\rho_a^h(\theta)\), normalized so that
\[
  L\rho_a(\theta)d\theta
\]
is the number of occupied rapidities of species \(a\) in \(d\theta\).  Define
the scattering kernel
\[
  \varphi_{ab}(\theta)
  =
  \frac{1}{2\pi}\frac{d}{d\theta}\delta_{ab}(\theta)
  =
  \frac{1}{2\pi\ii}\frac{d}{d\theta}\log S_{ab}(\theta).
\]
Differentiating the Bethe--Yang equations gives the density constraint
\[
  \rho_a(\theta)+\rho_a^h(\theta)
  =
  \frac{m_a}{2\pi}\cosh\theta
  +\sum_b\int_{\mathbb R}d\theta'\,
  \varphi_{ab}(\theta-\theta')\rho_b(\theta').
\]
This equation is purely kinematic: it counts available rapidity levels after
the scattering phase shifts have shifted the quantization conditions.

\section{Entropy and Free Energy}

For fixed total density \(\rho_a^t=\rho_a+\rho_a^h\), the number of
microscopic occupancy choices in an interval is
\[
  \binom{L\rho_a^t\,d\theta}{L\rho_a\,d\theta}.
\]
Stirling's formula gives the entropy density functional
\[
  s[\rho,\rho^h]
  =
  \sum_a\int d\theta\,
  \left[
  \rho_a^t\log\rho_a^t
  -\rho_a\log\rho_a
  -\rho_a^h\log\rho_a^h
  \right].
\]
At inverse temperature \(R\), the free energy per length is
\[
  f[\rho,\rho^h]
  =
  \sum_a\int d\theta\,m_a\cosh\theta\,\rho_a(\theta)
  -\frac1R s[\rho,\rho^h],
\]
subject to the density constraint.

\section{Pseudoenergy Equation}

Define the pseudoenergy by
\[
  \epsilon_a(\theta)=\log\frac{\rho_a^h(\theta)}{\rho_a(\theta)} .
\]
Varying the constrained free energy with respect to \(\rho_a\) gives
\[
  \epsilon_a(\theta)
  =
  Rm_a\cosh\theta
  -\sum_b\int_{\mathbb R}d\theta'\,
  \varphi_{ab}(\theta-\theta')
  \log\!\left(1+e^{-\epsilon_b(\theta')}\right).
\]
This is the thermodynamic Bethe ansatz equation.  The equilibrium free energy
is
\[
  f(R)
  =
  -\frac{1}{2\pi R}
  \sum_a m_a\int_{\mathbb R}d\theta\,
  \cosh\theta\,
  \log\!\left(1+e^{-\epsilon_a(\theta)}\right).
\]

\begin{proposition}[Variational derivation of TBA]
\label{prop:tba-variational-derivation}
For diagonal elastic scattering with smooth phase kernels and a thermodynamic
limit in which the Bethe--Yang root densities exist, the stationary point of
the constrained free-energy functional is exactly the pseudoenergy equation
above.
\end{proposition}

\begin{proof}
Work on the open cone where \(\rho_a>0\) and \(\rho_a^h>0\), and take compactly
supported smooth variations of the particle densities.  The density
constraint then determines the variation of the total density:
\[
  \delta\rho_c^t(\theta)
  =
  \sum_a\int d\theta'\,
  \varphi_{ca}(\theta-\theta')\delta\rho_a(\theta').
\]
Since \(\rho_a^h=\rho_a^t-\rho_a\), the corresponding hole-density variation
is \(\delta\rho_a^h=\delta\rho_a^t-\delta\rho_a\).
The entropy variation is
\[
  \delta s
  =
  \sum_a\int d\theta\,
  \left[
  \log\frac{\rho_a^h}{\rho_a}\,\delta\rho_a
  +
  \log\frac{\rho_a^t}{\rho_a^h}\,\delta\rho_a^t
  \right].
\]
Let
\[
  L_a(\theta)=\log\frac{\rho_a^t(\theta)}{\rho_a^h(\theta)}
  =
  \log(1+e^{-\epsilon_a(\theta)}).
\]
Substituting the constraint variation into \(\delta f=\delta e-R^{-1}\delta s\)
and collecting the coefficient of \(\delta\rho_a(\theta)\) gives
\[
  0
  =
  m_a\cosh\theta
  -\frac1R\epsilon_a(\theta)
  -\frac1R
  \sum_c\int d\theta'\,
  \varphi_{ca}(\theta'-\theta)L_c(\theta').
\]
With the diagonal-scattering convention
\(\varphi_{ca}(\theta'-\theta)=\varphi_{ac}(\theta-\theta')\), this is exactly
the displayed TBA equation.

It remains to derive the compact expression for the stationary free energy.
The entropy density identity
\[
  \rho_a^t\log\rho_a^t-\rho_a\log\rho_a-\rho_a^h\log\rho_a^h
  =
  \rho_a\,\epsilon_a+\rho_a^t L_a
\]
gives
\[
  Rf_\ast
  =
  \sum_a\int d\theta\,
  \rho_a(Rm_a\cosh\theta-\epsilon_a)
  -
  \sum_a\int d\theta\,\rho_a^t L_a .
\]
Use the stationary equation for the first parenthesis and the density
constraint for \(\rho_a^t\).  The convolution terms cancel after interchanging
species labels and integration variables, leaving only the one-particle
state-density term
\[
  Rf_\ast
  =
  -\sum_a\int d\theta\,
  \frac{m_a}{2\pi}\cosh\theta\,L_a(\theta),
\]
which is the stated formula for \(f(R)\).
\end{proof}

\section{Finite-Size Ground-State Energy}
\label{sec:tba-finite-size-ground-state}

Interchanging space and Euclidean time maps the thermal free energy on an
infinite line at inverse temperature \(R\) to the ground-state energy on a
circle of circumference \(R\), after the mirror-channel scattering data have
been identified.  In a relativistic theory the mirror dispersion is obtained
by a rapidity shift by \(\ii\pi/2\), so the same TBA equation gives
\[
  E_0(R)
  =
  -\frac{1}{2\pi}
  \sum_a m_a\int_{\mathbb R}d\theta\,
  \cosh\theta\,
  \log\!\left(1+e^{-\epsilon_a(\theta)}\right).
\]
The effective central charge extracted from the ultraviolet limit is
\[
  c_{\rm eff}(R)
  =
  -\frac{6R}{\pi}E_0(R),
  \qquad R\to0,
\]
provided the limit exists and the infrared bulk energy has been subtracted in
the declared normalization.

\section{Free Majorana Example and the Ultraviolet Central Charge}

The massive Ising field theory in its fermionic phase gives the minimal
diagonal example.  It has one neutral particle of mass \(m\).  The two-particle
scattering phase is the constant
\[
  S(\theta)=-1 .
\]
The derivative kernel is therefore zero:
\[
  \varphi(\theta)
  =
  \frac{1}{2\pi\ii}\frac{\dd}{\dd\theta}\log S(\theta)=0,
\]
after the constant branch of \(\log(-1)=\ii\pi\) has been fixed.  The TBA
equation reduces to
\[
  \epsilon(\theta)=r\cosh\theta,
  \qquad r=mR .
\]
The finite-size ground-state energy in the normalization of the preceding
section is
\begin{equation}
  E_0(R)
  =
  -\frac{m}{2\pi}
  \int_{-\infty}^{\infty}\dd\theta\,
  \cosh\theta\,
  \log\!\left(1+\ee^{-r\cosh\theta}\right).
  \label{eq:tba-free-majorana-ground-energy}
\end{equation}
The ultraviolet central charge follows from a direct scaling limit.  Put
\[
  x=\sinh\theta,\qquad \dd x=\cosh\theta\,\dd\theta .
\]
Then
\[
  I(r)
  :=
  \int_{-\infty}^{\infty}\dd\theta\,
  \cosh\theta\,
  \log(1+\ee^{-r\cosh\theta})
  =
  \int_{-\infty}^{\infty}\dd x\,
  \log(1+\ee^{-r\sqrt{1+x^2}}).
\]
Rescale \(y=rx\).  This gives
\[
  rI(r)
  =
  \int_{-\infty}^{\infty}\dd y\,
  \log(1+\ee^{-\sqrt{r^2+y^2}}).
\]
For \(0<r\leq1\) the integrand is bounded by
\(\log(1+\ee^{-|y|})\), which is integrable on \(\mathbb R\), and it
converges pointwise to \(\log(1+\ee^{-|y|})\).  Dominated convergence gives
\[
  \lim_{r\downarrow0}rI(r)
  =
  2\int_0^\infty\dd y\,\log(1+\ee^{-y}).
\]
Expanding \(\log(1+\ee^{-y})\) for \(y>0\) and integrating termwise gives
\[
  \sum_{k=1}^{\infty}\frac{(-1)^{k+1}}{k^2}
  =
  \int_0^\infty\dd y\,\log(1+\ee^{-y})
  =
  \frac{\pi^2}{12},
\]
Equation~\eqref{eq:tba-free-majorana-ground-energy} gives
\[
  E_0(R)
  =
  -\frac{\pi}{12R}+o(R^{-1}).
\]
Thus
\[
  c_{\rm eff}
  =
  \lim_{R\downarrow0}\left(-\frac{6R}{\pi}E_0(R)\right)
  =
  \frac12 .
\]
The result matches the central charge of the ultraviolet Majorana conformal
field theory, but the computation above uses only the declared finite-size
TBA specification and the normalization of \(E_0(R)\).

\section{Scaling Lee--Yang Example and the Effective Central Charge}

The first interacting one-particle TBA example is the massive scaling
Lee--Yang theory.  It is not a unitary Hilbert-space example; it is included
here because it is the minimal nontrivial diagonal bootstrap specification for which
the TBA ultraviolet limit can be carried to an exact number.  The ultraviolet
conformal theory is the nonunitary minimal model \(M(2,5)\), whose central
charge is \(c=-22/5\).  Its lowest conformal weight is \(-1/5\), so the
cylinder ground-state energy is governed by the effective central charge
\[
  c_{\mathrm{eff}}=c-24h_{\min}=\frac25 .
\]
The TBA computation below derives the same number from the scattering specification.

The diagonal one-particle amplitude is
\[
  S_{\mathrm{LY}}(\theta)
  =
  \frac{\sinh\theta+\ii\sin(\pi/3)}
       {\sinh\theta-\ii\sin(\pi/3)} .
\]
With the kernel convention of this chapter,
\begin{equation}
  \varphi_{\mathrm{LY}}(\theta)
  =
  \frac{1}{2\pi\ii}\frac{\dd}{\dd\theta}\log S_{\mathrm{LY}}(\theta)
  =
  -\frac{\sin(\pi/3)\cosh\theta}
        {\pi\left(\sinh^2\theta+\sin^2(\pi/3)\right)}
  =
  -\frac{2\sqrt3\cosh\theta}{\pi(2\cosh 2\theta+1)} .
  \label{eq:lee-yang-tba-kernel}
\end{equation}
It has total integral
\begin{equation}
  N_{\mathrm{LY}}
  =
  \int_{\mathbb R}\dd\theta\,\varphi_{\mathrm{LY}}(\theta)
  =
  -1.
  \label{eq:lee-yang-kernel-integral}
\end{equation}
Indeed, putting \(x=\sinh\theta\) and \(s=\sin(\pi/3)\) gives
\[
  N_{\mathrm{LY}}
  =
  -\frac{s}{\pi}
  \int_{-\infty}^{\infty}\frac{\dd x}{x^2+s^2}
  =
  -1.
\]
The TBA equation is therefore
\begin{equation}
  \epsilon(\theta)
  =
  r\cosh\theta
  -\int_{\mathbb R}\dd\theta'\,
  \varphi_{\mathrm{LY}}(\theta-\theta')
  \log(1+\ee^{-\epsilon(\theta')}).
  \label{eq:lee-yang-tba-equation}
\end{equation}

\begin{controlledapproximation}[Lee--Yang ultraviolet plateau]
\label{ca:lee-yang-tba-effective-central-charge}
Assume the standard one-plateau ultraviolet limit of
\eqref{eq:lee-yang-tba-equation}: as \(r\downarrow0\), there is a rapidity
region whose width tends to infinity in which
\(\epsilon(\theta)\) approaches a constant \(\epsilon_0\), while the kernel is
replaced by its integral \(N_{\mathrm{LY}}\).  Then
\[
  Y_0:=\ee^{\epsilon_0}=\phi_{\mathrm g}
  =
  \frac{1+\sqrt5}{2},
  \qquad
  c_{\mathrm{eff}}=\frac25 .
\]
\end{controlledapproximation}

The plateau equation is obtained from
\eqref{eq:lee-yang-tba-equation} by dropping the vanishing driving term and
replacing the convolution by \(N_{\mathrm{LY}}\):
\[
  \epsilon_0
  =
  -N_{\mathrm{LY}}\log(1+\ee^{-\epsilon_0})
  =
  \log(1+\ee^{-\epsilon_0}).
\]
Writing \(Y_0=\ee^{\epsilon_0}\) gives
\[
  Y_0=1+Y_0^{-1},
  \qquad
  Y_0^2=Y_0+1.
\]
The positive solution is the golden ratio \(\phi_{\mathrm g}\).

The ultraviolet TBA central charge is the one-species Rogers-dilogarithm
value
\[
  c_{\mathrm{eff}}
  =
  \frac{6}{\pi^2}
  L\!\left(\frac{1}{1+Y_0}\right)
  =
  \frac{6}{\pi^2}L(\phi_{\mathrm g}^{-2}).
\]
For \(0<x<1\), use
\[
  L(x)=\operatorname{Li}_2(x)+\frac12\log x\log(1-x).
\]
The two identities needed here are
\begin{equation}
  L(x)+L(1-x)=\frac{\pi^2}{6},
  \qquad
  L(x^2)=2L(x)-2L\!\left(\frac{x}{1+x}\right).
  \label{eq:rogers-euler-abel-identities}
\end{equation}
The first follows from Euler's dilogarithm identity.  The second follows by
differentiating the difference of the two sides, using
\(\operatorname{Li}_2'(x)=-\log(1-x)/x\), and evaluating at \(x=0\).
Set \(x=\phi_{\mathrm g}^{-1}\).  Then
\[
  1-x=x^2,\qquad
  \frac{x}{1+x}=x^2.
\]
The identities become
\[
  L(x)+L(x^2)=\frac{\pi^2}{6},
  \qquad
  L(x^2)=2L(x)-2L(x^2).
\]
Thus \(3L(x^2)=2L(x)\).  Solving the two linear equations gives
\[
  L(\phi_{\mathrm g}^{-2})=L(x^2)=\frac{\pi^2}{15}.
\]
Therefore \(c_{\mathrm{eff}}=6/15=2/5\).

\begin{remark}[What the example proves]
The calculation proves an internal consistency statement for the declared TBA
datum and its ultraviolet plateau asymptotics.  It is not by itself a
construction of the Lee--Yang local QFT, nor a derivation that the
finite-volume Hamiltonian spectrum of such a QFT is governed by
\eqref{eq:lee-yang-tba-equation}.  Those statements require the Hilbert-space
and locality input separated in the theorem boundary below.
\end{remark}

\paragraph{Status checkpoint: from scattering data to TBA.}
The TBA pseudoenergy equation has a different logical status from the
factorized scattering equations.  The scattering input supplies the kernel
\(\varphi=(2\pi\ii)^{-1}\partial_\theta\log S\).  To turn that kernel into a
thermodynamic statement about a local QFT one must also prove the
finite-volume Bethe--Yang description with controlled errors, convergence of
Bethe counting measures to root and hole densities, the microscopic entropy
of the allowed Bethe cells, existence and uniqueness of the free-energy
minimizer, and the channel identification that converts mirror free energy
into direct finite-size ground-state energy.  The free Majorana and Lee--Yang
computations above verify the internal finite algebra and ultraviolet
asymptotics of specific TBA data.  They do not by themselves prove that an
arbitrary local Hamiltonian with the same proposed scattering phase has that
finite-volume spectrum.

\section{Euler-Scale Generalized Hydrodynamics}
\label{sec:tba-generalized-hydrodynamics}

The equilibrium TBA datum also supplies the kinematic variables used in
generalized hydrodynamics.  In this chapter it is treated as a conditional
Euler-scale closure for inhomogeneous states whose local cells are assumed to
be described by thermodynamic Bethe root densities.

Let \(\rho_a(x,t,\theta)\) and \(\rho_a^h(x,t,\theta)\) be slowly varying
particle and hole densities, and set
\[
  \rho_a^t=\rho_a+\rho_a^h,
  \qquad
  n_a=\frac{\rho_a}{\rho_a^t}.
\]
The local density constraint is the pointwise version of the TBA counting
equation:
\begin{equation}
  \rho_a^t(x,t,\theta)
  =
  \frac{p'_a(\theta)}{2\pi}
  +
  \sum_b\int d\alpha\,
  \varphi_{ab}(\theta-\alpha)\rho_b(x,t,\alpha),
  \qquad
  p'_a(\theta)=m_a\cosh\theta .
  \label{eq:ghd-local-density-constraint}
\end{equation}
For any vector of one-particle functions \(h_a(\theta)\), define its dressing
by the linear equation
\begin{equation}
  h_a^{\rm dr}(\theta)
  =
  h_a(\theta)
  +
  \sum_b\int d\alpha\,
  \varphi_{ab}(\theta-\alpha)
  n_b(\alpha)h_b^{\rm dr}(\alpha).
  \label{eq:ghd-dressing-equation}
\end{equation}
The inverse of \(1-\varphi n\) is part of the hydrodynamic chart; if the
operator has a zero mode, this chart has left its domain of validity.
Applying \eqref{eq:ghd-dressing-equation} to \(p'_a\) gives
\[
  (p'_a)^{\rm dr}=2\pi\rho_a^t,
\]
which is precisely \eqref{eq:ghd-local-density-constraint}.  Applying it to
\[
  E'_a(\theta)=m_a\sinh\theta
\]
defines the Euler-scale effective velocity
\begin{equation}
  v_a^{\rm eff}(\theta)
  =
  \frac{(E'_a)^{\rm dr}(\theta)}
       {(p'_a)^{\rm dr}(\theta)} .
  \label{eq:ghd-effective-velocity}
\end{equation}

\begin{hypothesis}[Euler-scale GHD closure]
\label{hyp:euler-scale-ghd-closure}
Fix a diagonal massive TBA specification and a family of inhomogeneous finite-volume
states whose mesoscopic cells have thermodynamic root densities
\(\rho_a(x,t,\theta)\).  The Euler-scale generalized-hydrodynamic closure is
the assertion that, for every compact rapidity interval and every test
function in \(x,\theta\), the empirical quasiparticle measures converge in
the scaling limit to densities obeying
\begin{equation}
  \partial_t\rho_a(x,t,\theta)
  +
  \partial_x
  \left(
    v_a^{\rm eff}(x,t,\theta)\rho_a(x,t,\theta)
  \right)
  =
  0,
  \label{eq:ghd-root-density-continuity}
\end{equation}
with \(v_a^{\rm eff}\) defined from the local dressing equations
\eqref{eq:ghd-dressing-equation}--\eqref{eq:ghd-effective-velocity}.
\end{hypothesis}

This hypothesis has a direct conservation check.  For a local conserved
charge whose one-particle eigenvalue is \(h_a(\theta)\), define the Euler-cell
density and current by
\begin{equation}
  q_h(x,t)
  =
  \sum_a\int d\theta\,h_a(\theta)\rho_a(x,t,\theta),
  \qquad
  j_h(x,t)
  =
  \sum_a\int d\theta\,h_a(\theta)
  v_a^{\rm eff}(x,t,\theta)\rho_a(x,t,\theta).
  \label{eq:ghd-charge-current}
\end{equation}
Multiplying \eqref{eq:ghd-root-density-continuity} by \(h_a\), summing over
species, and integrating in rapidity gives
\[
  \partial_t q_h+\partial_xj_h=0
\]
whenever boundary terms vanish.  The closure is therefore compatible with the
infinite family of Bethe charges by construction.  The difficult assertion is
not this algebraic conservation check, but the scaling theorem that the
microscopic inhomogeneous QFT actually converges to
\eqref{eq:ghd-root-density-continuity}.

\begin{controlledapproximation}[Observable certificate for Euler GHD]
\label{ca:ghd-observable-reconstruction-certificate}
The formulas
\eqref{eq:ghd-root-density-continuity}--\eqref{eq:ghd-charge-current} are
hydrodynamic observables only after microscopic local densities and currents
have been matched to the local Bethe cell.  For a test function \(\chi(x)\)
and a conserved charge \(h\), define the Euler-cell predictions
\[
  Q_h^{\rm GHD}[\chi;t]
  =
  \int dx\,\chi(x)
  \sum_a\int d\theta\,h_a(\theta)\rho_a(x,t,\theta),
\]
and
\[
  J_h^{\rm GHD}[\chi;t]
  =
  \int dx\,\chi(x)
  \sum_a\int d\theta\,
  h_a(\theta)v_a^{\rm eff}(x,t,\theta)\rho_a(x,t,\theta).
\]
Let \(Q_{h,L}^{\rm mic}[\chi;t]\) and \(J_{h,L}^{\rm mic}[\chi;t]\) be the
corresponding smeared microscopic density and current operators in the finite
system.  A theorem-level GHD derivation must control the discrepancy
\begin{equation}
\begin{aligned}
  \mathcal E_{h,L}^{\rm GHD}[\chi;t]
  &=
  \left|
    \left\langle Q_{h,L}^{\rm mic}[\chi;t]\right\rangle
    -
    Q_h^{\rm GHD}[\chi;t]
  \right| \\
  &\quad+
  \left|
    \left\langle J_{h,L}^{\rm mic}[\chi;t]\right\rangle
    -
    J_h^{\rm GHD}[\chi;t]
  \right|.
\end{aligned}
  \label{eq:ghd-observable-error}
\end{equation}
The useful certificate is not a new conserved charge; it is a residual ledger:
\begin{equation}
  \mathcal E_{h,L}^{\rm GHD}[\chi;t]
  \le
  \epsilon_{\rm cell}
  +\epsilon_{\rm BY}
  +\epsilon_{\rm dress}
  +\epsilon_{\rm grad}
  +\epsilon_{\rm op}
  +\epsilon_{\rm diss}
  +\epsilon_{\rm break}.
  \label{eq:ghd-observable-residual-certificate}
\end{equation}
For the same test data and state family, \(\epsilon_{\rm cell}\) controls
replacement of the microscopic state in each mesoscopic box by a local
thermodynamic Bethe state;
\(\epsilon_{\rm BY}\) controls finite-volume Bethe--Yang/counting errors;
\(\epsilon_{\rm dress}\) controls stability of the inverse in
\eqref{eq:ghd-dressing-equation}; \(\epsilon_{\rm grad}\) controls gradients
discarded at Euler scale; \(\epsilon_{\rm op}\) controls the projection from
microscopic local density and current operators to the chosen one-particle
charge \(h_a(\theta)\); \(\epsilon_{\rm diss}\) controls diffusive, noise, and
memory corrections; and \(\epsilon_{\rm break}\) controls inelastic or
integrability-breaking channels.

This is the point at which GHD becomes physics rather than a reformulation of
TBA.  The bare velocity \(E'_a/p'_a\) is not the current velocity in an
interacting thermodynamic cell; the collision shifts already encoded in
\(\varphi\) change it to \(v_a^{\rm eff}\).  Conversely, even an exact solution
of \eqref{eq:ghd-root-density-continuity} certifies only the quasiparticle
density coordinate.  It does not by itself identify the microscopic current
operator, remove finite-volume recurrences, control gradients, or exclude
diffusive and inelastic corrections.
\end{controlledapproximation}

The formula for \(v^{\rm eff}\) is also the linearized counting formula for
a dressed wave packet.  Insert a single quasiparticle into a thermodynamic
cell and perturb the Bethe counting function.  The shift of available
rapidity levels is governed by the same kernel \(\varphi\) that appears in
\eqref{eq:ghd-local-density-constraint}.  The dressed momentum and dressed
energy gradients of the excitation are therefore \((p')^{\rm dr}\) and
\((E')^{\rm dr}\), and the group velocity of the dressed packet is their
ratio.  This is the kinematic content of
\eqref{eq:ghd-effective-velocity}; it does not prove equilibration to a
locally stationary Bethe state.

\paragraph{Hard-rod calibration.}
A useful finite algebraic laboratory is the classical hard-rod gas, which is
not a relativistic local QFT but tests the collision-shift logic.  Let rods
have length \(a>0\), velocity labels \(v_i\), and local velocity densities
\(\rho_i\) with \(a\rho_{\rm tot}<1\),
\[
  \rho_{\rm tot}=\sum_i\rho_i .
\]
The exact collision shift gives the effective-velocity equations
\begin{equation}
  v_i^{\rm eff}
  =
  v_i
  +
  a\sum_j\rho_j
  \left(v_i^{\rm eff}-v_j^{\rm eff}\right).
  \label{eq:hard-rod-effective-velocity-equation}
\end{equation}
Writing \(J=\sum_j\rho_jv_j^{\rm eff}\), this becomes
\[
  (1-a\rho_{\rm tot})v_i^{\rm eff}=v_i-aJ.
\]
Multiplying by \(\rho_i\) and summing over \(i\) gives
\[
  J=\sum_i\rho_iv_i .
\]
Hence
\begin{equation}
  v_i^{\rm eff}
  =
  \frac{v_i-a\sum_j\rho_jv_j}{1-a\rho_{\rm tot}} .
  \label{eq:hard-rod-effective-velocity-solution}
\end{equation}
The formula displays the same structure as
\eqref{eq:ghd-effective-velocity}: bare velocities are corrected by the
finite density of collision shifts, and the correction depends on the state.
In the quantum integrable case the scalar shift \(a\) is replaced by the
rapidity-dependent scattering kernel and the algebraic solution by the
dressing equation \eqref{eq:ghd-dressing-equation}.

\begin{openproblem}[GHD from local integrable QFT]
\label{op:ghd-from-local-integrable-qft}
Starting from a constructed local integrable QFT with factorized scattering,
derive the Euler-scale generalized-hydrodynamic closure
\eqref{eq:ghd-root-density-continuity} for inhomogeneous finite-energy
states.  The proof must identify the local thermodynamic Bethe state, control
finite-volume and mesoscopic errors, justify the dressing inverse, and show
that nonintegrable perturbations or inelastic channels are absent on the
Euler time scale under consideration.
\end{openproblem}

\section{Weak Integrability Breaking And Kinetic Cells}
\label{sec:tba-weak-breaking-kinetic-cells}

Perturbing an integrable QFT changes the role of the Bethe charges.  The
Hamiltonian still has exact energy, momentum, and internal-symmetry charges
declared by the perturbation, but the higher commuting charges of the
integrable point are generally lost.  A controlled kinetic description must
therefore specify which finite-cell transitions are allowed, which charges
they preserve exactly, and what entropy functional is monotone under the
coarse-grained collision dynamics.  These are finite algebraic requirements
before any limiting theorem from a local QFT is attempted.

\begin{definition}[Finite collision-cell Markov model]
\label{def:finite-collision-cell-datum}
A finite collision-cell Markov model consists of the following objects.

\begin{enumerate}
\item A finite set \(\Omega\) of mesoscopic cell states.
\item A strictly positive stationary weight \(\pi_\omega\) on \(\Omega\),
  normalized by \(\sum_{\omega\in\Omega}\pi_\omega=1\).
\item Transition rates \(k_{\omega\omega'}\ge0\), for
  \(\omega\ne\omega'\), satisfying detailed balance
  \begin{equation}
    \pi_\omega k_{\omega\omega'}
    =
    \pi_{\omega'} k_{\omega'\omega}.
    \label{eq:collision-cell-detailed-balance}
  \end{equation}
\item A finite family of exactly conserved cell observables
  \(Q^A:\Omega\to\mathbb R\), meaning that
  \(k_{\omega\omega'}\ne0\) implies \(Q^A(\omega)=Q^A(\omega')\) for every
  \(A\).
\item Occupation variables \(N_i:\Omega\to\mathbb Z_{\ge0}\), where \(i\)
  labels a finite rapidity/species bin.
\end{enumerate}

The cell probability \(P_\omega(T)\) evolves by
\begin{equation}
  \frac{dP_\omega}{dT}
  =
  \sum_{\omega'\ne\omega}
  \left(
    P_{\omega'}k_{\omega'\omega}
    -
    P_\omega k_{\omega\omega'}
  \right).
  \label{eq:collision-cell-master-equation}
\end{equation}
The projected occupation collision functional is
\begin{equation}
  C_i[P]
  =
  \sum_{\omega,\omega'}
  P_\omega k_{\omega\omega'}
  \left(N_i(\omega')-N_i(\omega)\right).
  \label{eq:collision-cell-projected-functional}
\end{equation}
If a conserved observable has the occupation form
\[
  Q^A(\omega)=Q^A_0+\sum_i h_i^A N_i(\omega),
\]
then the projected functional obeys
\begin{equation}
  \sum_i h_i^A C_i[P]=0 .
  \label{eq:collision-cell-projected-conservation}
\end{equation}
\end{definition}

\begin{lemma}[Finite collision-cell entropy identity]
\label{lem:finite-collision-cell-entropy-identity}
For a finite collision-cell Markov model, set
\[
  r_\omega=\frac{P_\omega}{\pi_\omega},
  \qquad
  \mathcal H(P|\pi)
  =
  \sum_{\omega\in\Omega}
  P_\omega\log\frac{P_\omega}{\pi_\omega}.
\]
Along \eqref{eq:collision-cell-master-equation},
\begin{equation}
  \frac{d}{dT}\mathcal H(P|\pi)
  =
  -
  \frac12
  \sum_{\omega,\omega'}
  \pi_\omega k_{\omega\omega'}
  \left(r_\omega-r_{\omega'}\right)
  \left(\log r_\omega-\log r_{\omega'}\right)
  \le0 .
  \label{eq:collision-cell-entropy-production}
\end{equation}
Moreover
\[
  \frac{d}{dT}\sum_\omega P_\omega Q^A(\omega)=0
\]
for every exactly conserved cell observable \(Q^A\).
\end{lemma}

\begin{proof}
The master equation conserves total probability, so the derivative of the
relative entropy is
\[
  \frac{d}{dT}\mathcal H(P|\pi)
  =
  \sum_\omega \dot P_\omega\log r_\omega .
\]
Insert \eqref{eq:collision-cell-master-equation} and exchange the dummy
indices in the gain term:
\[
  \frac{d}{dT}\mathcal H(P|\pi)
  =
  \sum_{\omega,\omega'}
  P_\omega k_{\omega\omega'}
  \left(\log r_{\omega'}-\log r_\omega\right).
\]
Using \(P_\omega=\pi_\omega r_\omega\) and detailed balance, the same sum
with \(\omega\) and \(\omega'\) exchanged is
\[
  \sum_{\omega,\omega'}
  \pi_\omega k_{\omega\omega'}r_{\omega'}
  \left(\log r_\omega-\log r_{\omega'}\right).
\]
Averaging the two expressions gives
\eqref{eq:collision-cell-entropy-production}.  The final sign follows from
monotonicity of the logarithm:
\[
  (x-y)(\log x-\log y)\ge0
  \qquad x,y>0 .
\]
For an exactly conserved cell observable, multiply the master equation by
\(Q^A(\omega)\), sum over \(\omega\), and symmetrize the result as
\[
  \frac{d}{dT}\sum_\omega P_\omega Q^A(\omega)
  =
  \sum_{\omega,\omega'}
  P_\omega k_{\omega\omega'}
  \left(Q^A(\omega')-Q^A(\omega)\right).
\]
Every nonzero rate has \(Q^A(\omega')=Q^A(\omega)\), hence the derivative
vanishes.  If \(Q^A\) is linear in the occupation variables, the same identity
is exactly \eqref{eq:collision-cell-projected-conservation}.
\end{proof}

\begin{controlledapproximation}[Weak-breaking kinetic scaling]
\label{ca:weak-breaking-kinetic-scaling}
Let \(H_0\) be an integrable Hamiltonian whose thermodynamic cells are
described by the TBA variables of Section~\ref{sec:tba-generalized-hydrodynamics},
and perturb it by a local operator,
\[
  H_\varepsilon=H_0+\varepsilon V,
  \qquad
  V=\int dx\,\mathcal O(x),
\]
with \(0<\varepsilon\ll1\).  A weak-breaking kinetic description is the
following scaling ansatz.  Choose a mesoscopic cell length \(\ell\) and a
macroscopic length \(X\) with microscopic scale \(\ll\ell\ll X\), pass to the
kinetic time \(T=\varepsilon^2t\), and assume that the local finite-cell
transition rates are obtained from the on-shell matrix elements of \(V\) after
the finite-volume limit, resonance averaging, and subtraction of disconnected
pieces have all been performed in a specified scheme.  The hydrodynamic
equation is then replaced by
\begin{equation}
  \partial_t\rho_i(x,t)
  +
  \partial_x\!\left(v_i^{\rm eff}[\rho](x,t)\rho_i(x,t)\right)
  =
  \varepsilon^2 C_i[P_{x,t}]
  +
  R_i^{(\varepsilon,\ell,X)}(x,t),
  \label{eq:weak-breaking-kinetic-ghd}
\end{equation}
where \(i=(a,\theta)\) denotes a species/rapidity bin after discretization,
\(P_{x,t}\) is the local cell probability whose occupation means give
\(\rho_i(x,t)\), and \(C_i\) is the projected collision functional
\eqref{eq:collision-cell-projected-functional}.  The remainder records
gradient corrections, memory effects, finite-cell correlations, and errors in
the kinetic limit.  The approximation is controlled only after one proves a
norm in which \(R_i^{(\varepsilon,\ell,X)}\to0\) under a declared joint limit
of \(\varepsilon,\ell,X\).
\end{controlledapproximation}

The finite-cell lemma identifies the part of the weak-breaking story that can
be checked before a local-QFT limit theorem has been proved.  If the
perturbation preserves energy and momentum, the collision graph must be
restricted to transitions
with the same energy and momentum labels; \eqref{eq:collision-cell-projected-conservation}
then gives the exact conservation laws of the kinetic equation.  If a higher
Bethe charge \(Q_h\) is not preserved by the perturbation, the same finite
formula displays its decay:
\[
  \frac{d}{dT}\langle Q_h\rangle
  =
  \sum_{\omega,\omega'}
  P_\omega k_{\omega\omega'}
  \left(Q_h(\omega')-Q_h(\omega)\right).
\]
Thus weak integrability breaking is not a modification of the dressing
formula alone.  It is a change in the collision graph of the mesoscopic cell,
with the remaining exact charges and the entropy production fixed before any
model-dependent scattering calculation is inserted.

\begin{openproblem}[Kinetic limit of weakly broken integrable QFT]
\label{op:weak-breaking-kinetic-limit}
Derive \eqref{eq:weak-breaking-kinetic-ghd} from a local integrable QFT
perturbed by a specified local operator.  The proof must construct the
finite-volume transition rates from regulated matrix elements, prove the
\(\varepsilon^{-2}\) kinetic time scaling, control secular terms and
near-degenerate resonances, identify the local stationary weights
\(\pi_\omega\), and show convergence of the remainder
\(R_i^{(\varepsilon,\ell,X)}\) in a norm strong enough to imply convergence
of the declared local observables.
\end{openproblem}

\section{Theorem Boundary}

The derivation above assumes existence of thermodynamic root densities,
validity of Bethe--Yang quantization up to errors negligible at large \(L\),
and uniqueness of the free-energy minimizer.  In concrete models these are
mathematical hypotheses to be proved or checked.  Bound states, non-diagonal
scattering, zero modes, massless flows, and weak integrability-breaking
collisions change the density system and must be included explicitly.

\begin{openproblem}[TBA from local QFT]
\label{op:tba-from-local-qft}
Derive the TBA equations directly from a constructed local QFT with
factorized scattering: prove Bethe--Yang quantization with controlled
finite-volume errors, establish thermodynamic root-density convergence, and
identify the finite-size ground-state energy with the mirror free energy.
\end{openproblem}
