\chapter{Perturbative Algebraic QFT On Curved Backgrounds}
\label{ch:curved-perturbative-algebraic-qft}

\ConventionsMixed

This chapter develops perturbative algebraic QFT on curved backgrounds as a
local construction of interacting fields from Hadamard two-point functions,
local functionals, controlled functional spaces, time-ordered products, and
causal factorization.  The wavefront-set condition called microcausality is
still essential: it is the local test that individual contractions with
Hadamard kernels are defined.  The complete algebraic foundation also
requires closure of the Peierls bracket, star product, and time-slice
homotopy, which in turn needs uniform control of how the derivative kernels
vary with the background field.  The equicausal class of
Hawkins--Rejzner--Visser supplies that control
\cite{HawkinsRejznerVisser2024Equicausal}.  The framework is perturbative,
but its locality, covariance, and functional-space conditions are
nonperturbative constraints on the allowed coefficients.

\section{Functional Spaces And Supports}

Let \(M\) be an oriented time-oriented globally hyperbolic Lorentzian
manifold and let
\[
  \mathcal E(M)=C^\infty(M,\mathbb R)
\]
be the classical configuration space of a real scalar field.  It is a
Frechet space with the usual seminorms obtained by taking suprema of
derivatives on compact subsets in coordinate charts.  A functional in this
chapter is a smooth map \(F:\mathcal E(M)\to\C\) in the Bastiani sense:
all iterated directional derivatives
\[
  F^{(n)}(\phi)(h_1,\ldots,h_n)
  =
  \left.
  \frac{\partial^n}{\partial t_1\cdots\partial t_n}
  F\!\left(\phi+\sum_{i=1}^nt_i h_i\right)
  \right|_{t_1=\cdots=t_n=0}
\]
exist and depend jointly continuously on \((\phi,h_1,\ldots,h_n)\).
When \(F^{(n)}(\phi)\) is distributional, the displayed multilinear form is
identified with a distributional density on \(M^n\) by
\[
  F^{(n)}(\phi)(h_1,\ldots,h_n)
  =
  \int_{M^n}
    F^{(n)}(\phi)(x_1,\ldots,x_n)
    h_1(x_1)\cdots h_n(x_n).
\]
This convention suppresses the volume density in the kernel notation.

The support of \(F\) is the closed set of points \(x\in M\) such that, for
every neighborhood \(U\ni x\), there are fields \(\phi\) and perturbations
\(h\in C_c^\infty(U)\) with
\[
  F(\phi+h)\ne F(\phi).
\]
Thus support is defined by actual sensitivity of the functional, not by a
chosen formula for it.

\begin{definition}[Local functional]
\label{def:paqft-local-functional}
A compactly supported smooth functional \(F\) is local if there are an
integer \(r\) and a compactly supported smooth density-valued function \(L\)
on the finite jet bundle \(J^rM\), polynomial in the highest derivatives in
the usual local-coordinate presentation, such that
\begin{equation}
  F(\phi)=\int_M L\bigl(x,j_x^r\phi\bigr),
  \label{eq:paqft-local-functional-jet-form}
\end{equation}
and such that \(F^{(n)}(\phi)\) is supported on the thin diagonal
\[
  \Delta_n=\{(x,\ldots,x):x\in M\}
\]
for every \(n\ge2\).  Equivalently, after integrations by parts,
\begin{equation}
  F^{(n)}(\phi)(x_1,\ldots,x_n)
  =
  \sum_{|\alpha_2|+\cdots+|\alpha_n|\le N}
  a_{\alpha_2\cdots\alpha_n}(x_1;\phi)
  \partial^{\alpha_2}\delta(x_1,x_2)\cdots
  \partial^{\alpha_n}\delta(x_1,x_n)
  \label{eq:paqft-local-functional-diagonal-kernel}
\end{equation}
in each coordinate chart, with coefficients depending locally and smoothly
on a finite jet of \(\phi\).
\end{definition}

The two descriptions in Definition~\ref{def:paqft-local-functional} are
both needed.  The jet formula is the geometric expression of locality; the
diagonal-support formula is the microlocal expression used in
renormalization.

\section{Microcausal And Equicausal Functionals}

\begin{definition}[Microcausal functional]
\label{def:paqft-microcausal-functional}
Let \(F:\mathcal E(M)\to\C\) be a compactly supported smooth functional
whose functional derivatives are compactly supported distributions,
\[
  F^{(n)}(\phi)\in\mathcal E'(M^n)
\]
for all \(n\).  The functional is microcausal if
\begin{equation}
  \operatorname{WF}(F^{(n)}(\phi))
  \cap
  \left((\overline V_+^*)^n\cup(\overline V_-^*)^n\right)
  =
  \varnothing .
  \label{eq:paqft-microcausal-cone-condition}
\end{equation}
Here \(\overline V_\pm^*\subset T^*M\setminus0\) are the closed future and
past causal covector cones, using the convention of
Chapter~\ref{ch:curved-microlocal-spectrum-condition}.  The vector space of
microcausal functionals is denoted \(\mathcal F_{\mu{\rm c}}(M)\), and the
subspace of local functionals is denoted \(\mathcal F_{\rm loc}(M)\).
\end{definition}

The exclusion in \eqref{eq:paqft-microcausal-cone-condition} is exactly the
condition needed to contract with Hadamard two-point functions.  A Hadamard
bidistribution has future covectors in one variable and past covectors in the
other.  If all covectors in \(F^{(n)}\) could point to the same closed time
cone, the product criterion of
Theorem~\ref{thm:microlocal-product-criterion} would fail when one contracts
the variables against the Hadamard kernel.

Microcausality is a pointwise condition in field space.  It says that each
kernel \(F^{(n)}(\phi)\) has an allowed wavefront set, hence that a displayed
contraction at that \(\phi\) is meaningful.  Closure demands the stronger
statement that the contracted expression is again a Bastiani-smooth
functional of \(\phi\), and that the class is stable under the Peierls
bracket, under the star product, and under the homotopy used in the time-slice
axiom.  The distinction is load bearing because a pAQFT observable is a
functional, not just a value of a distributional pairing at one background
field.

Let \(\mathcal E'_{\Gamma,K}(M^n)\) denote compactly supported
distributions on \(M^n\) with support in \(K\) and wavefront set contained in
the closed cone \(\Gamma\), equipped with the normal
H\"ormander topology described in
Chapter~\ref{ch:curved-microlocal-spectrum-condition}.  A compactly
supported smooth functional \(F\) is called equicausal if, for every
\(\phi_0\in\mathcal E(M)\) and every \(n\), there are a neighborhood
\(U\ni\phi_0\), a compact \(K\subset M^n\), and a closed cone
\[
  \Gamma_n
  \subset
  T^*M^n\setminus
  \left((\overline V_+^*)^n\cup(\overline V_-^*)^n\right)
\]
such that
\[
  F^{(n)}(\phi)\in\mathcal E'_{\Gamma_n,K}(M^n),
  \qquad \phi\in U,
\]
and the set
\(\{F^{(n)}(\phi):\phi\in U\}\) is equicontinuous in that normal topology.
The vector space of such functionals is denoted
\(\mathcal F_{\rm ec}(M)\).

The extra condition says that the singular directions stay uniformly away
from the forbidden causal cones, and that the seminorms controlling the
distributional kernels are bounded uniformly on a neighborhood in field
space.  This is the missing ingredient in the naive microcausal story:
one may perform the contraction and then differentiate, integrate over the
homotopy parameter, or compose with other kernels without losing smoothness.
The inclusions used below are
\[
  \mathcal F_{\rm loc}(M)
  \subset
  \mathcal F_{\rm ec}(M)
  \subset
  \mathcal F_{\mu{\rm c}}(M),
\]
and Wick polynomials and multilocal products generated from local
functionals lie in \(\mathcal F_{\rm ec}(M)\).

\begin{theorem}[Functional-space contract for curved pAQFT]
\label{thm:paqft-functional-space-contract}
On the equicausal class, the Peierls bracket and Hadamard star product close,
and the free-theory Koszul complex satisfies the time-slice axiom.  Local
functionals and Wick polynomials are equicausal.  The unrestricted
microcausal class \(\mathcal F_{\mu{\rm c}}(M)\) is used here only as a
wavefront-set pairing test unless a smaller polynomial, local, multilocal, or
equicausal subspace is explicitly named.
\end{theorem}

\begin{proof}
The product criterion gives pointwise distributional pairings from the
microcausal cone condition.  Equicausality adds uniform cone separation and
equicontinuity of the derivative kernels, so the same contractions define
smooth functionals and the Leibniz rules used in the Peierls and star-product
calculations are legitimate.  The same uniformity is preserved by the
chain-homotopy operations that implement the time-slice axiom.  Local
functionals have derivative kernels supported on diagonals with conormal
wavefront sets uniformly separated from the all-future and all-past cones;
Wick polynomials satisfy the same equicontinuity criterion.  These are the
closure results of Hawkins--Rejzner--Visser
\cite{HawkinsRejznerVisser2024Equicausal}.  Without the uniform step one has
only legal pointwise pairings, and there are microcausal examples whose Peierls
bracket fails to be a smooth functional.
\end{proof}

\section{Star Product}

Choose a Hadamard two-point function \(H\) whose antisymmetric part is
\[
  H(x,y)-H(y,x)=\ii\Delta(x,y),
\]
where \(\Delta=\Delta_{\rm ret}-\Delta_{\rm adv}\) is the causal propagator
of the Klein--Gordon operator.  For \(F,G\in\mathcal F_{\rm ec}(M)\),
define
\begin{equation}
  F\star_H G
  =
  \sum_{n=0}^{\infty}
  \frac{\hbar^n}{n!}
  \left\langle
    F^{(n)}(\phi), H^{\otimes n}G^{(n)}(\phi)
  \right\rangle .
  \label{eq:paqft-star-product-expanded}
\end{equation}
The \(n\)-th summand means
\begin{equation}
  \int_{M^{2n}}
    F^{(n)}(\phi)(x_1,\ldots,x_n)
    \prod_{j=1}^n H(x_j,y_j)
    G^{(n)}(\phi)(y_1,\ldots,y_n),
  \label{eq:paqft-star-product-kernel}
\end{equation}
with the integral interpreted as the pairing of distributions.  The
microcausal cone condition and the Hadamard wavefront set make each pairing
well-defined; equicausality is what makes the resulting expression a smooth
functional and keeps it in the same class.  For polynomial local or
multilocal functionals the sum is finite.  In general the product is a
formal series in the completed equicausal space
\(\mathcal F_{\rm ec}(M)[[\hbar]]\), order by order in \(\hbar\).

\paragraph{Associativity and commutator.}
The product \(\star_H\) is associative on the equicausal domain, with the
polynomial local subalgebra giving the finite calculation used in most
examples.  Its first-order commutator is
\begin{equation}
  [F,G]_{\star_H}
  =
  \ii\hbar\,\{F,G\}_{\rm P}
  +O(\hbar^2),
  \qquad
  \{F,G\}_{\rm P}
  =
  \left\langle F^{(1)},\Delta G^{(1)}\right\rangle ,
  \label{eq:paqft-peierls-first-order}
\end{equation}
where \(\{\,\cdot\,,\,\cdot\,\}_{\rm P}\) is the Peierls bracket of the free
field.
Let \(m\) denote pointwise multiplication and set
\[
  \Gamma_H=\left\langle H,\frac{\delta}{\delta\phi}\otimes
  \frac{\delta}{\delta\phi}\right\rangle
  \qquad\hbox{on}\qquad
  \mathcal F\otimes\mathcal F .
\]
Then
\[
  F\star_HG=m\circ e^{\hbar\Gamma_H}(F\otimes G).
\]
For three factors let \(\Gamma_{ij}\) contract the \(i\)-th and \(j\)-th
tensor factors.  Functional derivatives with respect to different tensor
factors commute, so
\[
  [\Gamma_{12},\Gamma_{13}]
  =
  [\Gamma_{12},\Gamma_{23}]
  =
  [\Gamma_{13},\Gamma_{23}]
  =
  0 .
\]
Both \((F\star_HG)\star_HK\) and \(F\star_H(G\star_HK)\) are therefore equal
to
\[
  m_3\circ
  \exp\bigl(\hbar(\Gamma_{12}+\Gamma_{13}+\Gamma_{23})\bigr)
  (F\otimes G\otimes K),
\]
where \(m_3(F\otimes G\otimes K)=FGK\).  This proves associativity.

The order-\(\hbar\) part of the commutator is
\[
  \hbar\left\langle F^{(1)},(H-H^{\rm op})G^{(1)}\right\rangle
  =
  \ii\hbar\left\langle F^{(1)},\Delta G^{(1)}\right\rangle ,
\]
which is the free Peierls bracket.

\section{Change Of Hadamard Function}

\begin{proposition}[Change of Hadamard function]
\label{prop:paqft-hadamard-change-isomorphism}
If \(H'\) is another Hadamard two-point function, then
\[
  w=H'-H
\]
is smooth.  Define
\[
  \alpha_w
  =
  \exp\!\left(
    \frac{\hbar}{2}
    \left\langle w,\frac{\delta^2}{\delta\phi^2}\right\rangle
  \right).
\]
Then
\begin{equation}
  \alpha_w(F\star_H G)=\alpha_w(F)\star_{H'}\alpha_w(G).
  \label{eq:paqft-hadamard-change-isomorphism}
\end{equation}
Thus the abstract algebra is independent of the particular Hadamard
parametrix up to a canonical isomorphism determined by the smooth difference.
\end{proposition}

\begin{proof}
Let
\[
  \Delta_w=\left\langle w,\frac{\delta^2}{\delta\phi^2}\right\rangle .
\]
Because \(w\) is smooth, \(\Delta_w\) maps the polynomial local subalgebra,
and more generally the equicausal domain order by order, to itself.  The
second-order operator \(\Delta_w\) satisfies
\begin{equation}
  \Delta_w(FG)
  =
  (\Delta_wF)G+F(\Delta_wG)
  +
  2\left\langle F^{(1)},wG^{(1)}\right\rangle .
  \label{eq:paqft-laplacian-product-rule}
\end{equation}
Equivalently,
\begin{equation}
  e^{\frac{\hbar}{2}\Delta_w}\circ m
  =
  m\circ e^{\hbar\Gamma_w}
    \left(e^{\frac{\hbar}{2}\Delta_w}\otimes
          e^{\frac{\hbar}{2}\Delta_w}\right),
  \label{eq:paqft-alpha-coproduct-identity}
\end{equation}
where \(\Gamma_w\) is the cross-contraction with kernel \(w\).  Therefore
\[
\begin{aligned}
  \alpha_w(F\star_HG)
  &=
  e^{\frac{\hbar}{2}\Delta_w}
  m e^{\hbar\Gamma_H}(F\otimes G)\\
  &=
  m e^{\hbar\Gamma_w}
  \left(e^{\frac{\hbar}{2}\Delta_w}\otimes
        e^{\frac{\hbar}{2}\Delta_w}\right)
  e^{\hbar\Gamma_H}(F\otimes G)\\
  &=
  m e^{\hbar\Gamma_{H+w}}(\alpha_wF\otimes\alpha_wG)
  =
  \alpha_w(F)\star_{H'}\alpha_w(G).
\end{aligned}
\]
\end{proof}

\section{Time-Ordered Products}

Let \(H_F=H+\ii\Delta_{\rm adv}\) denote the Feynman bidistribution
associated with \(H\).  Away from coincident points, the naive time-ordered
product is obtained by replacing \(H\) in the contraction exponential by
\(H_F\).  The problem is that \(H_F\) has a wavefront set whose products are
not defined on diagonals.  Thus time ordering is not a pointwise operation on
local fields; it is an extension problem for distributions.

A system of time-ordered products is a family of maps
\begin{equation}
  T_n:\mathcal F_{\rm loc}(M)^{\otimes n}
  \longrightarrow
  \mathcal F_{\rm ec}(M)[[\hbar]],
  \qquad n\ge1,
  \label{eq:paqft-time-ordered-products-maps}
\end{equation}
with \(T_1=\operatorname{id}\), symmetric in the \(n\) entries, equal to the
Feynman contraction formula away from partial diagonals, and satisfying:
\[
\begin{gathered}
  \text{local covariance},\qquad
  \text{causal factorization},\\
  \text{field independence},\\
  \text{scaling degree bounds},\qquad
  \text{unitarity and Ward identities}.
\end{gathered}
\]
These words are constraints, not decoration.  Local covariance means natural
dependence on the background metric and couplings.  Field independence means
that differentiating \(T_n\) with respect to the background field inserts the
time-ordered derivative of the input local functionals.  Scaling degree
bounds control the ultraviolet order of singularities.  Ward identities are
the equations imposed by symmetries; for gauge theories they are properly
encoded by the BV quantum master equation.

The domain in \eqref{eq:paqft-time-ordered-products-maps} is deliberately
local.  Epstein--Glaser renormalization constructs the extension of products
of local interactions and local fields; local and multilocal functionals are
equicausal, so the resulting time-ordered products live in the same closed
functional-space lane used by the star product.  A general time-ordering
operator on arbitrary equicausal functionals is not assumed here; extending
the Epstein--Glaser construction beyond the local/multilocal input class is
an additional open problem, not a prerequisite for the worked local
interacting examples below.

\begin{theorem}[Extension with fixed scaling degree]
\label{thm:paqft-extension-scaling-degree}
Let \(t^0\in\mathcal D'(\mathbb R^N\setminus\{0\})\) have finite scaling
degree
\[
  \operatorname{sd}(t^0)
  =
  \inf\left\{\delta:
    \lambda^\delta t^0(\lambda x)\to0
    \text{ in }\mathcal D'(\mathbb R^N\setminus\{0\})
    \text{ as }\lambda\to0^+
  \right\}.
\]
Then \(t^0\) has an extension \(t\in\mathcal D'(\mathbb R^N)\) with
\(\operatorname{sd}(t)=\operatorname{sd}(t^0)\).  If
\(\operatorname{sd}(t^0)<N\), this extension is unique.  If
\(\operatorname{sd}(t^0)\ge N\), two such extensions differ by
\begin{equation}
  \sum_{|\alpha|\le \operatorname{sd}(t^0)-N}
  c_\alpha\,\partial^\alpha\delta_0 ,
  \label{eq:paqft-extension-ambiguity}
\end{equation}
with the upper bound understood as its integer part.
\end{theorem}

\begin{proof}
Choose a test function \(\rho\in C_c^\infty(\mathbb R^N)\) equal to \(1\)
near the origin.  For a test function \(f\), subtract its Taylor polynomial
at the origin through degree
\[
  r=\lfloor\operatorname{sd}(t^0)\rfloor-N
\]
when \(r\ge0\):
\begin{equation}
  f_R(x)=
  f(x)-
  \rho(x)\sum_{|\alpha|\le r}\frac{x^\alpha}{\alpha!}
  (\partial^\alpha f)(0).
  \label{eq:paqft-taylor-subtracted-test-function}
\end{equation}
The subtracted test function vanishes to order \(r+1\) at the origin.  The
definition of scaling degree then implies that \(t^0(f_R)\) is finite and
independent of how \(f_R\) is represented away from the origin.  Set
\begin{equation}
  t(f)=t^0(f_R)+\sum_{|\alpha|\le r}c_\alpha(\partial^\alpha f)(0).
  \label{eq:paqft-extension-by-taylor-subtraction}
\end{equation}
The constants \(c_\alpha\) are arbitrary finite renormalization constants.
The subtraction raises the vanishing order of the test function precisely
enough to preserve the original scaling degree, and the additional
\(\partial^\alpha\delta_0\) terms have scaling degree \(N+|\alpha|\le
\operatorname{sd}(t^0)\).

If \(r<0\), no Taylor coefficient may be added without increasing the scaling
degree.  An extension is then forced by continuity from
\(\mathbb R^N\setminus\{0\}\), hence unique.  In the case \(r\ge0\), any two
extensions agree away from the origin, so their difference is a distribution
supported at \(0\).  Every distribution supported at one point is a finite
linear combination of derivatives of the delta distribution, and the scaling
degree bound permits exactly the derivatives in
\eqref{eq:paqft-extension-ambiguity}.
\end{proof}

In pAQFT the role of the origin in
Theorem~\ref{thm:paqft-extension-scaling-degree} is played by a partial
diagonal in \(M^n\).  The finite ambiguity is therefore a finite sum of local
curvature, mass, coupling, and field polynomials multiplying derivatives of
delta distributions supported on that diagonal.  These are contact terms
created by the definition of time-ordered products, not a separate physical
process.

\section{Causal Factorization}

If the supports of \(F_1,\ldots,F_k\) lie nowhere in the past of the supports
of \(F_{k+1},\ldots,F_n\), causal factorization states
\[
  T_n(F_1,\ldots,F_n)
  =
  T_k(F_1,\ldots,F_k)\star
  T_{n-k}(F_{k+1},\ldots,F_n).
\]
This identity determines \(T_n\) outside the total diagonal from lower orders.
Renormalization is precisely the choice of extension to the total and partial
diagonals compatible with the listed constraints.

The induction is local in configuration space.  On a configuration
\((x_1,\ldots,x_n)\) not lying on the total diagonal, global hyperbolicity
provides a Cauchy time function and one may separate the points into a later
cluster and an earlier cluster.  Causal factorization expresses \(T_n\) there
as a \(\star_H\)-product of already constructed lower-order time-ordered
products.  On overlaps of such cluster decompositions the lower-order
factorization identities and associativity of \(\star_H\) give the same
distribution.  Hence the only new datum at order \(n\) is the extension
across the total diagonal, subject to covariance, scaling degree, unitarity,
and any Ward identities.

\section{Interacting Fields}

Let \(V\) be an interaction functional with compact time support.  The
relative S-matrix is
\begin{equation}
  S_V(F)=S(V)^{-1}\star S(V+F),
  \label{eq:paqft-curved-relative-s-matrix}
\end{equation}
where
\[
  S(F)=T\exp\!\left(\frac{\ii}{\hbar}F\right).
\]
The interacting field associated to a local functional \(F\) is the
Bogoliubov derivative
\[
  R_V(F)
  =
  \left.\frac{\hbar}{\ii}\frac{d}{d\lambda}
  S_V(\lambda F)\right|_{\lambda=0}.
\]
Changing the time cutoff of \(V\) outside the causal domain of \(F\) gives
isomorphic interacting algebras by causal factorization.  The useful form is
the generalized Bogoliubov factorization identity for relative S-matrices:
if \(\operatorname{supp}P\cap J_-(\operatorname{supp}Q)=\varnothing\), then
\begin{equation}
  S_V(P+G+Q)
  =
  S_V(P+G)\star S_V(G)^{-1}\star S_V(G+Q).
  \label{eq:paqft-bogoliubov-factorization-relative}
\end{equation}
This is the algebraic adiabatic-limit mechanism of the local interacting net,
as in the construction reviewed by Fredenhagen--Rejzner
\cite[Sec.~4]{FredenhagenRejzner2015PerturbativeConstruction}.  Two
one-sided consequences are load bearing.  If a cutoff change \(C\) is outside
the causal past of \(F\),
\[
  \operatorname{supp}C\cap J_-(\operatorname{supp}F)=\varnothing,
\]
then
\begin{equation}
  S_{V+C}(F)=S_V(F).
  \label{eq:paqft-cutoff-outside-past}
\end{equation}
Indeed, \eqref{eq:paqft-bogoliubov-factorization-relative} gives
\(S_V(C+F)=S_V(C)\star S_V(F)\), and the factors \(S_V(C)\) cancel in the
definition of \(S_{V+C}(F)\).  If \(C\) is outside the causal future of \(F\),
\[
  \operatorname{supp}C\cap J_+(\operatorname{supp}F)=\varnothing,
\]
then
\begin{equation}
  S_{V+C}(F)
  =
  S_V(C)^{-1}\star S_V(F)\star S_V(C).
  \label{eq:paqft-cutoff-outside-future}
\end{equation}
Here \eqref{eq:paqft-bogoliubov-factorization-relative} gives
\(S_V(F+C)=S_V(F)\star S_V(C)\).  Notice that the conjugating element is the
relative cocycle \(S_V(C)\), not the raw formal S-matrix \(S(C)\); \(C\) may
overlap the background interaction \(V\).

\begin{proposition}[Causal support of the Bogoliubov field]
\label{prop:paqft-bogoliubov-causal-support}
If two interaction functionals \(V\) and \(V'\) agree on a neighborhood of
the causal hull \(J(\operatorname{supp}F)\), then the corresponding
Bogoliubov fields \(R_V(F)\) and \(R_{V'}(F)\) are identified by the
canonical relative \(S\)-matrix isomorphism of the two cutoff theories.
\end{proposition}

\begin{proof}
Let \(C=V'-V\).  By hypothesis \(C\) vanishes on a neighborhood of
\(J(\operatorname{supp}F)\).  Choose two Cauchy surfaces, one to the past of
\(\operatorname{supp}F\) and one to its future, both lying inside the region
where \(V=V'\) near the causal hull.  A smooth partition of unity in the
complement of that hull decomposes
\[
  C=A+B,
\]
with
\[
  \operatorname{supp}A\cap J_-(\operatorname{supp}F)=\varnothing,
  \qquad
  \operatorname{supp}B\cap J_+(\operatorname{supp}F)=\varnothing .
\]
The future piece is removed first, using
\eqref{eq:paqft-cutoff-outside-past} with background \(V+B\):
\[
  S_{V+B+A}(\lambda F)=S_{V+B}(\lambda F).
\]
The past piece is then handled by
\eqref{eq:paqft-cutoff-outside-future}:
\begin{equation}
  S_{V'}(\lambda F)
  =
  S_V(B)^{-1}\star S_V(\lambda F)\star S_V(B).
  \label{eq:paqft-cutoff-transition-generator}
\end{equation}
Thus the cutoff change is implemented on the local generators by the
insertion-independent relative-cocycle isomorphism
\[
  \alpha_{V,B}(X)=S_V(B)^{-1}\star X\star S_V(B).
\]
Differentiating \eqref{eq:paqft-cutoff-transition-generator} at
\(\lambda=0\) gives
\[
  R_{V'}(F)=\alpha_{V,B}(R_V(F)).
\]

The auxiliary decomposition does not add data.  If a third cutoff \(W\) is
inserted between \(V\) and \(V'\), the relative S-matrices obey the cocycle
identity
\[
  S_V(D+E)=S_V(D)\star S_{V+D}(E),
\]
which is \eqref{eq:paqft-bogoliubov-factorization-relative} with no
observable insertion.  Hence the transition \(V\to V'\) is the composition of
the transitions \(V\to W\) and \(W\to V'\) on all generators supported in the
chosen local region.  This is exactly the evaluation-map construction of the
algebraic adiabatic limit: local algebras are represented by any compact
cutoff agreeing with the physical interaction near
\(J_+(\mathcal O)\cap J_-(\mathcal O)\), and the transition maps above are
the canonical identifications between such representatives.
\end{proof}

\section{Renormalization Group}

Two choices of time-ordered products \(T\) and \(\widetilde T\) are related
by maps \(Z\) on local functionals:
\[
  \widetilde S(F)=S(Z(F)).
\]
The map \(Z\) is local, begins with \(Z(F)=F+O(F^2)\), preserves the constant
term, and has coefficients supported on coincident points.  In curved
backgrounds these coefficients include curvature terms.  This is the
perturbative algebraic form of local counterterm freedom.

More explicitly, \(Z\) is a formal diffeomorphism of the space of local
functionals:
\begin{equation}
  Z(F)
  =
  F+
  \sum_{n\ge2}\frac1{n!}Z_n(F^{\otimes n}),
  \label{eq:paqft-z-expansion}
\end{equation}
where each \(Z_n\) is symmetric, local, and supported on the thin diagonal in
the \(n\)-fold product.  The defining identity
\(\widetilde S(F)=S(Z(F))\) implies that changing the extension of
time-ordered products is equivalent to changing the local interaction
functional by finite local terms.  In a curved background those terms are
not limited to flat-space monomials: covariance permits terms built from the
metric, curvature, masses, background couplings, and their covariant
derivatives, constrained by dimension, symmetries, and field independence.

\section{Worked Interacting Scalar Coordinate}
\label{sec:paqft-worked-interacting-scalar-coordinate}

The previous sections describe the abstract perturbative construction.  A
curved-background interacting example must also say which local coordinates
are being held fixed and what observable changes when those coordinates are
changed.  Consider a real scalar with compact spacetime cutoff \(g\), and
write the quartic interaction in a Hadamard Wick coordinate \(H\) as
\begin{equation}
  V_{\lambda,H}(g)
  =
  \int_M \dd\operatorname{vol}_g\,
  g(x)\,\frac{\lambda}{4!}\Phi_H^4(x).
  \label{eq:paqft-lambda-phi-four-interaction-coordinate}
\end{equation}
Let \(H'=H+w\), with \(w\) smooth, and write
\[
  w_\Delta(x)=w(x,x).
\]
The smooth-change isomorphism
Proposition~\ref{prop:paqft-hadamard-change-isomorphism} gives the local
diagonal identities
\begin{equation}
\begin{aligned}
  \alpha_w\bigl(\Phi_H^2(x)\bigr)
  &=
  \Phi_{H'}^2(x)+\hbar w_\Delta(x),\\
  \alpha_w\bigl(\Phi_H^4(x)\bigr)
  &=
  \Phi_{H'}^4(x)
  +6\hbar w_\Delta(x)\Phi_{H'}^2(x)
  +3\hbar^2w_\Delta(x)^2 .
\end{aligned}
  \label{eq:paqft-wick-power-scheme-transport}
\end{equation}
Thus the same abstract quartic interaction, transported to the \(H'\)
coordinate, is
\begin{equation}
\begin{aligned}
  \alpha_w V_{\lambda,H}(g)
  =
  \int_M\dd\operatorname{vol}_g\,g(x)
  \left[
    \frac{\lambda}{4!}\Phi_{H'}^4
    +\frac{\hbar\lambda}{4}w_\Delta\Phi_{H'}^2
    +\frac{\hbar^2\lambda}{8}w_\Delta^2
  \right](x).
\end{aligned}
  \label{eq:paqft-phi-four-scheme-shift}
\end{equation}
Equations \eqref{eq:paqft-wick-power-scheme-transport} and
\eqref{eq:paqft-phi-four-scheme-shift} are algebra-coordinate statements.
They do not by themselves produce a physical expectation-value shift.  If
\(\omega_H\) is a state on the \(H\)-presentation and \(\omega_{H'}\) is the
transported state on the \(H'\)-presentation, defined by
\[
  \omega_{H'}(\alpha_w A)=\omega_H(A),
\]
then
\[
  \omega_{H'}\!\left(\alpha_w V_{\lambda,H}(g)\right)
  =
  \omega_H\!\left(V_{\lambda,H}(g)\right).
\]
This is the same abstract observable in two Hadamard coordinates.

A finite local Wick-renormalization change is a different comparison.  It is
not an arbitrary smooth difference of two Hadamard two-point functions; local
covariance restricts the diagonal scalar inserted into Wick powers.  In four
dimensions, for the dimension-two part, write
\begin{equation}
  c(x)=a_m m^2+a_R R(x).
  \label{eq:paqft-local-wick-finite-renormalization}
\end{equation}
Changing the Wick prescription by this local scalar uses the same polynomial
combinatorics as \eqref{eq:paqft-wick-power-scheme-transport}, with
\(w_\Delta\) replaced by \(c\), but it compares two local covariant
composite-field prescriptions in a fixed physical state.  The quartic
interaction then acquires the finite local terms
\[
  \int_M\dd\operatorname{vol}_g\,g(x)
  \left[
    \frac{\hbar\lambda}{4}c\Phi_H^2
    +
    \frac{\hbar^2\lambda}{8}c^2
  \right](x).
\]
The quadratic term is the finite coordinate shift
\begin{equation}
  \delta m^2=\frac{\hbar\lambda}{2}a_m m^2,
  \qquad
  \delta\xi=\frac{\hbar\lambda}{2}a_R
  \label{eq:paqft-phi-four-mass-curvature-shifts}
\end{equation}
in the local scalar operator
\(\frac12(m^2+\xi R)\Phi^2\).  The last term in
\eqref{eq:paqft-phi-four-scheme-shift} is a finite shift of the purely
geometric source functional, producing \(m^4\), \(m^2R\), and \(R^2\)
coordinates when \eqref{eq:paqft-local-wick-finite-renormalization} is used.
Those coordinates are precisely the ones that reappear in the stress-tensor
and gravitational EFT bookkeeping of
Chapters~\ref{ch:point-splitting-stress-tensor-renormalization}
and~\ref{ch:curved-semiclassical-backreaction}.

This paragraph records only the Wick-coordinate subfamily of the finite
renormalization freedom.  A complete interacting stress-tensor construction
also fixes the time-ordered products used in the Bogoliubov map, the
interaction-dependent composite-field counterterms in \(\Phi^4\), and the
independent conserved local tensors allowed in
Proposition~\ref{prop:finite-local-freedom-scalar-stress-tensor}.  The
Hollands--Wald conservation theorem is a condition on that complete
construction: the chosen time-ordered products and contact terms must make the
interacting stress tensor conserved, rather than relying on the potential
insertion or the Wick-square shift by themselves
\cite{HollandsWald2001LocalWick,HollandsWald2005Conservation,HollandsWald2015QFCS}.

The local Wick-square comparison has a visible c-number output only in this
fixed-state, changed-prescription sense.  If
\[
  F_f=\int_M\dd\operatorname{vol}_g\, f(x)\Phi_H^2(x),
  \qquad f\in C_c^\infty(M),
\]
and the second Wick-square prescription is
\[
  F_f^{(c)}
  =
  F_f+\hbar\int_M\dd\operatorname{vol}_g\,f(x)c(x),
\]
then
\begin{equation}
  \omega_H\!\left(F_f^{(c)}\right)-\omega_H(F_f)
  =
  \hbar\int_M\dd\operatorname{vol}_g\,f(x)c(x).
  \label{eq:paqft-wick-square-observable-shift}
\end{equation}
The right side is state independent because it is a local renormalization
coordinate, not particle production and not the transport of a state between
two algebra presentations.  A physical renormalization condition fixes
\(a_m,a_R\) or fixes the value of
\eqref{eq:paqft-wick-square-observable-shift} on selected backgrounds and
states.  Without that condition, the number is a coordinate of the composite
operator, not an invariant prediction.

\paragraph{Derivation.}
For a local polynomial at one point, the smooth-change isomorphism acts by
\[
  \alpha_w
  =
  \exp\left(\frac{\hbar}{2}w_\Delta
  \frac{\partial^2}{\partial\phi^2}\right).
\]
Applying this operator to \(\phi^2\) and \(\phi^4\) gives
\eqref{eq:paqft-wick-power-scheme-transport}.  Multiplication by
\(\lambda/4!\) gives the coordinate-transport identity
\eqref{eq:paqft-phi-four-scheme-shift}.  The local Wick-renormalization
comparison uses the same finite polynomial identity with the locally
covariant scalar \(c\) in place of the generic smooth diagonal
\(w_\Delta\); the identities
\eqref{eq:paqft-phi-four-mass-curvature-shifts} follow by comparing the
coefficient of \(\frac12(m^2+\xi R)\Phi^2\).  Finally, applying the first
line of \eqref{eq:paqft-wick-power-scheme-transport} with
\(w_\Delta\) replaced by \(c\) inside \(F_f\) gives
\eqref{eq:paqft-wick-square-observable-shift}.

\section{Worked Local-Coupling Ward Ledger}
\label{sec:paqft-local-coupling-ward-ledger}

The compact spacetime cutoff in
\eqref{eq:paqft-lambda-phi-four-interaction-coordinate} is a construction
device, not a harmless decoration on a stress source.  To keep the metric
notation separate in this paragraph, write the switching function as
\(\chi\) and define the local interaction-density insertion by
\begin{equation}
  \mathcal O_{\lambda,H}(x)
  :=
  \frac{\delta V_{\lambda,H}(\chi)}{\delta\chi(x)}
  =
  \frac{\lambda}{4!}\Phi_H^4(x)
  \label{eq:paqft-local-coupling-density}
\end{equation}
before the additional finite composite-field counterterms are chosen.  Let
\(\mathcal W[g_{\mu\nu},\chi]\) be the renormalized source functional in a
fixed state and prescription, with variations
\[
  \delta\mathcal W
  =
  -\frac12
  \int_M\dd\operatorname{vol}_g\,
  \langle T_{\mu\nu}^{(\chi)}\rangle\delta g^{\mu\nu}
  +
  \int_M\dd\operatorname{vol}_g\,
  \langle \mathcal O_\lambda^{(\chi)}\rangle\delta\chi .
\]
If the renormalized time-ordered products and composite fields are chosen
covariantly and there is no diffeomorphism anomaly, simultaneous pullback of
the metric, state data, and background coupling by a compactly supported flow
gives
\begin{equation}
  \nabla^\mu\langle T_{\mu\nu}^{(\chi)}\rangle
  =
  \langle \mathcal O_\lambda^{(\chi)}\rangle\,\nabla_\nu\chi
  \quad
  \text{modulo field-equation ideals and contact terms.}
  \label{eq:paqft-local-coupling-ward-identity}
\end{equation}
The sign is fixed by the displayed variation convention for
\(\mathcal W\).  Indeed, for a vector field \(X\),
\[
  0
  =
  \delta_X\mathcal W
  =
  \int_M\dd\operatorname{vol}_g\,
  \langle T_{\mu\nu}^{(\chi)}\rangle\nabla^\mu X^\nu
  +
  \int_M\dd\operatorname{vol}_g\,
  \langle\mathcal O_\lambda^{(\chi)}\rangle X^\nu\nabla_\nu\chi ,
\]
and integration by parts gives
\eqref{eq:paqft-local-coupling-ward-identity}.  Thus a compactly supported
interaction has a controlled Ward balance, not a conserved stress tensor by
itself.  The switching edge is an external background that injects or removes
four-momentum from the quantum system.  On an observation region where
\(\nabla\chi=0\), the right side vanishes and the algebraic adiabatic
isomorphism can compare different cutoffs outside the causal domain.  On the
edge region, however, one may not feed the partial interaction stress into a
semiclassical Einstein equation as though it were conserved.

The identity also organizes finite renormalization.  A change of the
renormalized density
\[
  \mathcal O_\lambda^{(\chi)}
  \longmapsto
  \mathcal O_\lambda^{(\chi)}+\delta\mathcal O_{\rm loc}
\]
changes the right side of
\eqref{eq:paqft-local-coupling-ward-identity} by
\(\langle\delta\mathcal O_{\rm loc}\rangle\nabla_\nu\chi\).  The stress tensor
counterterms and contact extensions must change in the same prescription so
that the Ward balance remains true.  Shifting the composite operator while
leaving the stress tensor source fixed, or shifting the potential stress while
forgetting the coupling-source insertion, is not a new physical prediction;
it is an inconsistent comparison of schemes.

For the \(\lambda\Phi^4\) example in a centered quasifree reference state,
the first displayed expectation of
\eqref{eq:paqft-local-coupling-density} is
\[
  \left\langle\mathcal O_{\lambda,H}(x)\right\rangle_\omega
  =
  \frac{\lambda}{8}\Sigma_{\omega,H}(x)^2 .
\]
Therefore the finite-switching Ward source sampled by a compact vector field
is
\begin{equation}
  \int_M\dd\operatorname{vol}_g\,
  X^\nu
  \frac{\lambda}{8}\Sigma_{\omega,H}^2\nabla_\nu\chi
  \label{eq:paqft-local-coupling-finite-switch-source}
\end{equation}
inside this retained state coordinate.  A constant replacement of
\(\Sigma_{\omega,H}\) or an averaged replacement of \(\nabla\chi\) is not an
equivalent source unless the corresponding residual is shown to be small in
the tested response norm.  This is the pAQFT origin of the Ward-clean source
requirement used in
Chapter~\ref{ch:curved-semiclassical-backreaction}: the physical
backreaction input is the full scheme-fixed interacting stress tensor, or a
retained source with an explicit Ward residual and error budget, not the
potential insertion alone.

\paragraph{Finite check.}
The companion calculation represents \(\chi\), the Wick-square coordinate,
and compact vector tests on four finite cells.  It verifies the discrete
version of the balance
\[
  \delta_X\mathcal W_{\rm metric}
  +
  \delta_X\mathcal W_{\rm coupling}=0,
  \qquad
  \delta_X\mathcal W_{\rm coupling}
  =
  \sum_i w_iX_i
  \frac{\lambda}{8}\Sigma_i^2(\nabla\chi)_i ,
\]
checks that a constant switching function gives no coupling Ward source, and
rejects three common shortcuts: declaring a compactly switched interaction
conserved, replacing a nonconstant density by its average, and applying a
finite Wick-density shift without transporting the Ward source in the same
scheme.

\section{Worked Interacting Tadpole Response}
\label{sec:paqft-worked-interacting-tadpole-response}

The preceding example is still a coordinate calculation unless it is tied to
an observable response.  A minimal physics-facing output is the one-loop local
tadpole part of the interacting two-point response.  Let \(P\) be the free
Klein--Gordon operator, let \(E_{\rm ret}^0\) and \(E_{\rm adv}^0\) be its
retarded and advanced Green operators, and let \(g\) be the compact
interaction cutoff in
\eqref{eq:paqft-lambda-phi-four-interaction-coordinate}.  Fix a quasifree
Hadamard state \(\omega\) and a Wick prescription \(H\).  The smooth
state-dependent Wick square is
\[
  \Sigma_{\omega,H}(x)=\omega\!\left(\Phi_H^2(x)\right).
\]
In a stationary or ultrastatic region where \(g=1\) and
\(\Sigma_{\omega,H}\) is slowly varying, this quantity is the local tadpole
density seen by a propagating quantum.  The retained local one-loop
self-energy coordinate is
\begin{equation}
  \mathsf M_{\rm tad}(x)
  =
  \frac{\lambda}{2}\,g(x)\Sigma_{\omega,H}(x).
  \label{eq:paqft-tadpole-mass-coordinate}
\end{equation}
The factor \(1/2\) is the quartic vertex factor
\(\lambda/4!\) times the \(12\) Wick contractions that choose two legs to
form the local tadpole and two legs to attach to the external two-point
line.  Equivalently, differentiating the interaction twice gives
\(\lambda\Phi_H^2/2\), and the quasifree state replaces the Wick square by
\(\Sigma_{\omega,H}\).

The corresponding first Born correction to the retarded two-point response
is not a new local functional; it is the inverse-operator identity
\[
  (P+\mathsf M_{\rm tad})^{-1}_{\rm ret}
  =
  E_{\rm ret}^0
  -
  E_{\rm ret}^0\mathsf M_{\rm tad}E_{\rm ret}^0
  +O(\lambda^2)
\]
read in the retarded algebra.  For compact test functions \(f,h\), formal
self-adjointness gives the smeared response shift
\begin{equation}
\begin{aligned}
  \delta_\lambda G_{\rm ret}(f,h)
  &=
  -\int_M\dd\operatorname{vol}_g\,
    (E_{\rm adv}^0f)(x)\,
    \mathsf M_{\rm tad}(x)\,
    (E_{\rm ret}^0h)(x)\\
  &=
  -\frac{\lambda}{2}
  \int_M\dd\operatorname{vol}_g\,
    g(x)\Sigma_{\omega,H}(x)
    (E_{\rm adv}^0f)(x)(E_{\rm ret}^0h)(x).
\end{aligned}
  \label{eq:paqft-tadpole-born-response}
\end{equation}
This is a genuine response statement: the observable is the retarded
two-point function, the state input is the renormalized local Wick square,
and the output is the first-order shift of propagation through the region in
which the interaction is turned on.  If \(\Sigma_{\omega,H}\) is not
constant, it must remain inside the integral.  Pulling it out as an average
loses the spatial overlap between the tadpole density and the two Green
functions and is not an equivalent amplitude.

The same comparison also shows where finite Wick renormalization enters a
physical response.  Replacing the Wick-square prescription by
\(\Phi_H^2+\hbar c\), with \(c=a_m m^2+a_R R\), changes the retained tadpole
coordinate by
\begin{equation}
  \mathsf M_{\rm tad}^{(c)}(x)-\mathsf M_{\rm tad}(x)
  =
  \frac{\hbar\lambda}{2}\,g(x)c(x),
  \label{eq:paqft-tadpole-scheme-response-shift}
\end{equation}
and hence shifts \eqref{eq:paqft-tadpole-born-response} by the same Born
formula with \(\Sigma_{\omega,H}\) replaced by \(\hbar c\).  A
renormalization condition on a two-point response, pole mass, or local
curvature expansion fixes this finite coordinate.  Without that condition,
\eqref{eq:paqft-tadpole-scheme-response-shift} is scheme bookkeeping, not a
separate physical prediction.

The restriction to \(\mathsf M_{\rm tad}\) is deliberate.  The full
interacting two-point function also contains nonlocal self-energy terms,
derivative and curvature contact terms, cutoff-edge corrections in the
adiabatic limit, state corrections, and higher-loop contributions.  Those
terms contain the subtler physical content.  Equation
\eqref{eq:paqft-tadpole-born-response} is the retained local part for which
the coefficient, sign, state dependence, and scheme dependence can be
checked without pretending that the whole interacting propagator has been
computed.

\section{Worked Nonlocal Self-Energy Response}
\label{sec:paqft-worked-nonlocal-self-energy-response}

The local tadpole is only the first term in the interacting two-point
response.  The first contribution at order \(\lambda^2\) that cannot be replaced
by a local mass, curvature, or wavefunction coordinate is the sunset
self-energy.  Work away from the diagonal \(x=y\), keep the same compact
cutoff \(g\), and write the greater and lesser two-point functions of the
chosen quasifree Hadamard state as
\[
  W_\omega^>(x,y)=\omega(\Phi(x)\Phi(y)),
  \qquad
  W_\omega^<(x,y)=W_\omega^>(y,x).
\]
For a Hermitian real scalar field,
\(W_\omega^<(x,y)=\overline{W_\omega^>(x,y)}\) as a distributional boundary
value away from the diagonal.  Let
\(\Theta_{\rm ret}(x,y)\) denote the retarded support factor, supported where
\(x\in J^+(y)\).  Following the closed-time-path convention
\[
  G^{\rm R}=-\ii\,\Theta_{\rm ret}(G^>-G^<)
\]
from Chapter~\ref{ch:thermal-schwinger-keldysh}, retardedizing the cubic
Wightman product supplies the factor \(-\ii\).  The off-diagonal sunset kernel
inserted into the inverse two-point operator is
\begin{equation}
  \Pi_{\rm sun,off}^{\rm ret}(x,y)
  =
  -\frac{\lambda^2}{6}\,
  g(x)g(y)\,
  \Theta_{\rm ret}(x,y)
  \left[-\ii
  \left(
    W_\omega^>(x,y)^3-W_\omega^<(x,y)^3
  \right)\right],
  \qquad x\ne y .
  \label{eq:paqft-sunset-retarded-kernel-offdiagonal}
\end{equation}
For Hermitian Wightman data the parenthesized cubic difference is purely
imaginary, so the displayed kernel is real on real test functions.  Omitting
the \(-\ii\) factor would turn this time-domain inverse-kernel insertion into
an imaginary source for a real retarded response.
This is the first nonlocal part of the two-point response: the value at
\((x,y)\) remembers propagation through three internal state contractions
between two interaction points.  A replacement by a diagonal scalar such as
\(\Sigma_{\omega,H}(x)\) would erase the physical channel that gives
dissipation, thresholds, and memory.

The response of the retarded two-point function is the same inverse-operator
identity used in \eqref{eq:paqft-tadpole-born-response}, but now with a
bilocal kernel:
\begin{equation}
  \delta_{\lambda^2}G_{\rm ret}^{\rm sun}(f,h)
  =
  -\int_{M^2}
  (E_{\rm adv}^0f)(x)\,
  \Pi_{\rm sun}^{\rm ret}(x,y)\,
  (E_{\rm ret}^0h)(y)\,
  \dd\operatorname{vol}_g(x)\dd\operatorname{vol}_g(y).
  \label{eq:paqft-sunset-born-response}
\end{equation}
The kernel \(\Pi_{\rm sun}^{\rm ret}\) in
\eqref{eq:paqft-sunset-born-response} includes
\eqref{eq:paqft-sunset-retarded-kernel-offdiagonal} together with its extension
across \(x=y\).  That Epstein--Glaser extension has local freedom:
\begin{equation}
  \Pi_{\rm sun}^{\rm ret}
  =
  \operatorname{Ext}_{\rm EG}
  \Pi_{\rm sun,off}^{\rm ret}
  +
  \left[
    b_ZP_x+b_m m^2+b_R R(x)+\cdots
  \right]\delta(x,y).
  \label{eq:paqft-sunset-local-extension-freedom}
\end{equation}
The displayed local terms are representative members of the same finite
two-point counterterm family that fixes field normalization, mass, curvature
coupling, and higher local curvature coordinates at the chosen dimension.
They are required data for promoting
\eqref{eq:paqft-sunset-born-response} from a separated-point kernel to a
renormalized interacting propagator.

\paragraph{Coefficient and causality check.}
The coefficient \(1/6\) in
\eqref{eq:paqft-sunset-retarded-kernel-offdiagonal} is a real combinatorial
statement before any curved-spacetime extension problem appears.  In the
second cumulant of
\(\lambda\Phi^4/4!\), choose one external field at each vertex: \(4\cdot4\)
choices.  The remaining three fields at \(x\) and three fields at \(y\) can
be paired across the vertices in \(3!\) ways.  The cumulant prefactor gives
\[
  -\frac12\left(\frac{\lambda}{4!}\right)^2(4\cdot4)(3!)
  =
  -\frac{\lambda^2}{12}
\]
as the coefficient of the bilocal quadratic action
\(\int\varphi(x)g(x)W(x,y)^3g(y)\varphi(y)\).  The two-point inverse kernel
is twice the quadratic-action coefficient, hence
\(-\lambda^2/6\).  Passing to retarded response replaces the symmetric
Feynman product by the closed-time-path causal difference
\(-\ii\Theta_{\rm ret}\bigl((W_\omega^>)^3-(W_\omega^<)^3\bigr)\).  The finite
check attached to this chapter verifies this off-diagonal coefficient, the
retarded support, the \(-\ii\) reality factor on Hermitian Wightman samples, and
the Born sign.  The continuum distributional extension and the physical
renormalization condition fixing
\eqref{eq:paqft-sunset-local-extension-freedom} remain additional data.

\section{Gauge Theories And The BV Condition}

The preceding scalar construction is not a foundation for gauge theory by
itself.  A gauge-fixed perturbative algebra contains unphysical fields,
ghosts, antifields, and a differential.  Consistency is the statement that
the interacting quantum BV differential squares to zero, or equivalently
that the renormalized time-ordered products can be chosen so that the
quantum master equation holds:
\begin{equation}
  \frac12(S_{\rm BV},S_{\rm BV})-\ii\hbar\,\Delta_{\rm BV}S_{\rm BV}=0
  \label{eq:paqft-bv-qme-curved}
\end{equation}
in the renormalized algebra.  Here \((\,\cdot\,,\,\cdot\,)\) is the BV
antibracket and \(\Delta_{\rm BV}\) denotes the renormalized second-order
operator measuring the short-distance contraction of fields with antifields.
The equation is not a slogan for gauge invariance.  It is the algebraic
condition that the interacting BRST operator is a differential and that the
physical algebra is its cohomology.  If the obstruction to solving
\eqref{eq:paqft-bv-qme-curved} is a nontrivial local cohomology class, the
theory has an anomaly in the chosen background and field content.

\begin{openproblem}[Nonperturbative curved-background comparison]
\label{op:nonperturbative-paqft-comparison}
Relate perturbative algebraic QFT on curved backgrounds to nonperturbative
constructions by comparing local algebras, Hadamard states, stress tensors,
renormalized composite fields, anomaly terms, and scaling limits in examples
where both sides are available.
\end{openproblem}
