\chapter{Background Gauge Fields and Index Theory}
\label{ch:curved-background-gauge-fields-index-theory}

\ConventionsEuclidean

Background gauge fields are not auxiliary decorations.  They are sources for
global currents, they specify the bundles in which charged fields take
values, and for chiral fermions they determine both zero-mode selection
rules and anomaly lines.  This chapter develops the index-theoretic part of
that structure with explicit conventions.  The analytic input is the index
of a coupled Dirac operator; the QFT input is the Berezin integral over its
zero modes and the determinant line of its nonzero modes.
The source traces for the heat-kernel/index-theory line used here are
Seeley, Gilkey, and Atiyah--Singer
\cite{Seeley1967,Gilkey1984,AtiyahSinger1968}; the derivations below keep the
local Clifford and Chern-Weil conventions explicit because these are the
conventions later used in anomaly formulae.

\section{Spinors, Bundles, and Curvature Conventions}

Let \(M^{2n}\) be a closed oriented Riemannian spin manifold.  The complex
spinor bundle splits into chiral pieces
\[
  S=S^+\oplus S^- .
\]
Let \(E\to M\) be a Hermitian vector bundle with unitary connection
\(\nabla^E\).  In the characteristic-class formulas of this chapter the
curvature \(F_E=(\nabla^E)^2\) is skew-Hermitian, so
\[
  \operatorname{ch}(E)
  =
  \operatorname{tr}_E\exp\!\left(\frac{\ii F_E}{2\pi}\right)
\]
is a real even cohomology class.  If a physics convention uses Hermitian
generators and Hermitian curvature \(F_{\rm phys}\), then
\[
  F_E=-\ii F_{\rm phys}
\]
in this formula.  All traces in \(\operatorname{ch}(E)\) are traces in the
fermion representation, not necessarily the trace used in the Yang-Mills
kinetic term.

The coupled Dirac operator is
\[
  D_A:\Gamma(S\otimes E)\longrightarrow\Gamma(S\otimes E),
  \qquad
  D_A=\gamma^\mu\nabla_\mu^{S\otimes E}.
\]
With respect to chirality,
\[
  D_A=
  \begin{pmatrix}
  0 & D_A^-\\
  D_A^+ & 0
  \end{pmatrix},
  \qquad
  D_A^+:\Gamma(S^+\otimes E)\to\Gamma(S^-\otimes E).
\]
The analytic index is
\begin{equation}
  \operatorname{Ind}(D_A^+)
  =
  \dim\ker D_A^+-\dim\ker D_A^- .
  \label{eq:index-analytic-definition}
\end{equation}
Because \(M\) is compact and \(D_A^+\) is elliptic, this is a finite integer.

\section{McKean--Singer Supertrace}

Let \(\Gamma\) denote the chirality operator, acting by \(+1\) on \(S^+\) and
by \(-1\) on \(S^-\).  The heat operator \(e^{-tD_A^2}\) is trace class for
\(t>0\).

\begin{proposition}[McKean--Singer identity]
\label{prop:index-mckean-singer}
For every \(t>0\),
\begin{equation}
  \operatorname{Str}\,e^{-tD_A^2}
  :=
  \operatorname{Tr}\!\left(\Gamma e^{-tD_A^2}\right)
  =
  \operatorname{Ind}(D_A^+).
  \label{eq:index-mckean-singer}
\end{equation}
\end{proposition}

\begin{proof}
Because \(M\) is compact and \(D_A\) is elliptic and self-adjoint, \(D_A^2\)
has discrete spectrum with finite-dimensional eigenspaces and
\(e^{-tD_A^2}\) is trace class for \(t>0\).  The chirality operator anticommutes
with \(D_A\), so \(D_A\) exchanges \(S^+\otimes E\) and \(S^-\otimes E\).  The
nonzero spectrum of \(D_A^2\) on \(S^+\otimes E\) is paired with the nonzero
spectrum on \(S^-\otimes E\).  Indeed, if
\(\psi_+\in\Gamma(S^+\otimes E)\) satisfies
\[
  D_A^-D_A^+\psi_+=\lambda\psi_+,\qquad \lambda>0,
\]
then \(D_A^+\psi_+\neq0\) and
\[
  D_A^+D_A^-(D_A^+\psi_+)=\lambda(D_A^+\psi_+).
\]
The inverse map on the \(\lambda\)-eigenspace is \(\lambda^{-1}D_A^-\), so the
pairing preserves multiplicities.  Hence the positive eigenvalue
contributions to
\[
  \operatorname{Tr}_{S^+}e^{-tD_A^-D_A^+}
  -
  \operatorname{Tr}_{S^-}e^{-tD_A^+D_A^-}
\]
cancel term by term, with absolute convergence supplied by the heat trace.
The zero eigenspaces are precisely \(\ker D_A^+\) and \(\ker D_A^-\).  Their
contribution is independent of \(t\) and equals
\[
  \dim\ker D_A^+-\dim\ker D_A^-=\operatorname{Ind}(D_A^+).
\]
\end{proof}

This proof is finite-dimensional after elliptic spectral theory supplies the
discrete spectrum and heat-kernel trace class property.  It is the analytic
bridge between zero-mode counting and local curvature densities.

\section{Lichnerowicz Formula and Heat-Kernel Scaling}

The local index theorem is a statement about the short-time asymptotics of
the heat kernel of \(D_A^2\).  We spell out the conventions because the same
heat-kernel coefficients also enter point-splitting stress tensors and trace
anomalies.

Let \(c(e^a)=\gamma^a\) denote Clifford multiplication by an orthonormal
coframe.  We use
\[
  \gamma^{ab}:=\frac12[\gamma^a,\gamma^b],
  \qquad
  F^E=\frac12F^E_{ab}e^a\wedge e^b .
\]
The curvature of the spin connection is
\[
  [\nabla_a^S,\nabla_b^S]
  =
  \frac14 R_{ab cd}\gamma^{cd}.
\]

\begin{lemma}[Lichnerowicz formula]
\label{lem:index-lichnerowicz}
For the coupled Dirac operator \(D_A=\gamma^a\nabla_a^{S\otimes E}\),
\begin{equation}
  D_A^2
  =
  (\nabla^{S\otimes E})^\ast\nabla^{S\otimes E}
  +\frac14{\rm Scal}
  +\frac12\gamma^{ab}F^E_{ab}.
  \label{eq:index-lichnerowicz}
\end{equation}
\end{lemma}

\begin{proof}
At a chosen point \(p\), take an oriented orthonormal frame whose Levi-Civita
connection coefficients vanish at \(p\).  Write \(\nabla\) for the coupled
connection on \(S\otimes E\).  Since the antisymmetric part of
\(\gamma^a\gamma^b\) is \(\gamma^{ab}\),
\[
\begin{aligned}
  D_A^2
  &=
  \gamma^a\gamma^b\nabla_a\nabla_b                                      \\
  &=
  g^{ab}\nabla_a\nabla_b
  +\gamma^{ab}\nabla_a\nabla_b                                          \\
  &=
  g^{ab}\nabla_a\nabla_b
  +\frac12\gamma^{ab}[\nabla_a,\nabla_b].
\end{aligned}
\]
Restoring the Christoffel term away from the chosen normal frame gives the
connection Laplacian \((\nabla^{S\otimes E})^\ast\nabla^{S\otimes E}\) in the
sign convention of this chapter.  The curvature of the coupled connection is
\[
  [\nabla_a,\nabla_b]
  =
  \frac14R_{ab cd}\gamma^{cd}+F^E_{ab}.
\]
The \(E\)-bundle curvature therefore contributes
\[
  \frac12\gamma^{ab}F^E_{ab}.
\]
The spin-curvature contribution is
\[
  \frac12\gamma^{ab}\frac14R_{ab cd}\gamma^{cd}.
\]
It remains to evaluate the Clifford contraction.  With
\(\gamma^{abcd}:=\gamma^{[a}\gamma^b\gamma^c\gamma^{d]}\), the Clifford
relations give
\[
  \gamma^{ab}\gamma^{cd}
  =
  \gamma^{abcd}
  +g^{bc}\gamma^{ad}
  -g^{ac}\gamma^{bd}
  -g^{bd}\gamma^{ac}
  +g^{ad}\gamma^{bc}
  +g^{ac}g^{bd}
  -g^{ad}g^{bc}.
\]
The coefficient of \(\gamma^{abcd}\) is the total antisymmetrization of
\(R_{abcd}\), hence vanishes by the first Bianchi identity
\(R_{[abc]d}=0\).  The terms with one remaining \(\gamma^{pq}\) reduce to
Ricci contractions, for example \(R_{ab cd}g^{bc}\gamma^{ad}\), and vanish
because the Ricci tensor is symmetric while \(\gamma^{ad}\) is antisymmetric.
The scalar terms give
\[
  R_{ab cd}(g^{ac}g^{bd}-g^{ad}g^{bc})
  =
  2\,{\rm Scal}
\]
with the curvature and Clifford conventions stated above.  Thus the
spin-curvature piece is
\[
  \frac18\,R_{ab cd}\gamma^{ab}\gamma^{cd}
  =
  \frac14{\rm Scal}.
\]
Combining the connection Laplacian, spin curvature, and bundle curvature gives
\eqref{eq:index-lichnerowicz}.
\end{proof}

\begin{definition}[Laplace-type operator and heat coefficients]
\label{def:index-laplace-type-heat-coefficients}
A Laplace-type operator on a Hermitian vector bundle \(V\to M^d\) is written
in the form
\begin{equation}
  P=-(g^{\mu\nu}\nabla_\mu\nabla_\nu+\mathcal E),
  \label{eq:index-laplace-type-convention}
\end{equation}
where \(\nabla\) is a connection on \(V\) and
\(\mathcal E\in\Gamma({\rm End}\,V)\).  Its curvature is
\(\Omega_{\mu\nu}=[\nabla_\mu,\nabla_\nu]\).  The local heat coefficients
\(a_{2r}(P;x)\) are defined by
\begin{equation}
  {\rm Tr}\!\left(f e^{-tP}\right)
  \sim
  (4\pi t)^{-d/2}
  \sum_{r\geq0}t^r
  \int_M f(x)\,{\rm tr}_V a_{2r}(P;x)\,d{\rm vol}_g
  \label{eq:index-heat-coeff-definition}
\end{equation}
for every test function \(f\).
\end{definition}

\begin{quotedtheorem}[First Seeley--DeWitt coefficients]
\label{qthm:index-seeley-dewitt-coefficients}
\emph{Source trace.}  The displayed normalization follows the
Seeley--DeWitt/Gilkey heat-kernel convention
\cite{Seeley1967,Gilkey1984}.
For the convention \eqref{eq:index-laplace-type-convention},
\begin{equation}
  a_0={\bf 1}_V,
  \qquad
  a_2=\mathcal E+\frac16{\rm Scal}\,{\bf 1}_V,
  \label{eq:index-a0-a2}
\end{equation}
and
\begin{equation}
\begin{aligned}
  a_4
  =
  \frac1{360}\Big[
  &\left(5{\rm Scal}^2-2{\rm Ric}_{\mu\nu}{\rm Ric}^{\mu\nu}
  +2R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma}\right){\bf 1}_V \\
  &+60{\rm Scal}\,\mathcal E
  +180\mathcal E^2
  +30\Omega_{\mu\nu}\Omega^{\mu\nu} \\
  &+12\nabla^2{\rm Scal}\,{\bf 1}_V
  +60\nabla^2\mathcal E
  \Big].
  \label{eq:index-a4-seeley-dewitt}
\end{aligned}
\end{equation}
On a closed manifold the last line integrates to zero after taking a trace.
\end{quotedtheorem}

\paragraph{Parametrix origin.}
Work in a normal convex neighborhood and write Synge's world function as
\(\sigma(x,y)\).  The heat kernel parametrix has the form
\begin{equation}
  K_t(x,y)
  \sim
  (4\pi t)^{-d/2}e^{-\sigma(x,y)/2t}\Delta(x,y)^{1/2}
  \sum_{r\geq0}t^r b_r(x,y),
  \qquad b_0(x,x)={\bf 1}_V,
  \label{eq:index-heat-parametrix}
\end{equation}
where \(\Delta\) is the van Vleck determinant and \(b_r\) is parallel
transported in the bundle indices at leading order.  Inserting
\eqref{eq:index-heat-parametrix} into \((\partial_t+P_x)K_t=0\) and comparing
powers of \(t\) gives the transport recursion
\begin{equation}
  (\sigma^{;\mu}\nabla_\mu+r)b_r
  =
  -\Delta^{-1/2}
  P_x\!\left(\Delta^{1/2}b_{r-1}\right),
  \qquad r\geq1.
  \label{eq:index-heat-transport-recursion}
\end{equation}
The coincidence limits needed below are
\begin{equation}
  [\sigma^{;\mu}]=0,\qquad
  [\sigma^{;\mu}{}_{\nu}]=\delta^\mu{}_\nu,\qquad
  [\nabla_\mu\nabla_\nu\Delta^{1/2}]
  =\frac16{\rm Ric}_{\mu\nu},
  \label{eq:index-world-function-coincidence}
\end{equation}
with brackets denoting \(y\to x\).  Setting \(r=1\) in
\eqref{eq:index-heat-transport-recursion} immediately gives
\[
  [b_1]=\mathcal E+\frac16{\rm Scal}\,{\bf 1}_V .
  \]
For \(r=2\), differentiate the transport equation twice, use
\eqref{eq:index-world-function-coincidence}, the commutator
\([\nabla_\mu,\nabla_\nu]=\Omega_{\mu\nu}\), and commute covariant
derivatives until all third and fourth derivatives of \(\sigma\) and
\(\Delta^{1/2}\) are expressed in curvature.  The resulting coincidence
limit is precisely \eqref{eq:index-a4-seeley-dewitt}.  No field equation is
used; the calculation is the universal local recursion for a Laplace-type
operator.

\section{Local Index Theorem}

Let \(R\) denote the Riemannian curvature two-form of \(TM\).  The
\(\widehat A\)-class is
\[
  \widehat A(TM)
  =
  \det\nolimits^{1/2}
  \left(\frac{R/4\pi}{\sinh(R/4\pi)}\right)
  =
  1-\frac{p_1(TM)}{24}
  +\frac{7p_1(TM)^2-4p_2(TM)}{5760}+\cdots .
\]
The theorem below is the heat-kernel/Getzler route to the local
Atiyah--Singer index formula \cite{AtiyahSinger1968,Gilkey1984}; its proof is
included to fix the curvature and chirality normalization used by the anomaly
polynomial later in the chapter.

\begin{theorem}[Local index density by Getzler rescaling]
\label{thm:index-atiyah-singer-local}
For the chiral Dirac operator coupled to \(E\),
\begin{equation}
  \operatorname{Ind}(D_A^+)
  =
  \int_M
  \left[\widehat A(TM)\operatorname{ch}(E)\right]_{2n}.
  \label{eq:index-atiyah-singer}
\end{equation}
Equivalently, the local heat-kernel density of the supertrace has
\[
  \lim_{t\to0^+}{\rm str}\,K_t(x,x)\,d{\rm vol}_g
  =
  \left[\widehat A(TM)\operatorname{ch}(E)\right]_{2n}.
\]
\end{theorem}

\begin{proof}
The McKean--Singer identity reduces the index to the small-\(t\) limit of the
local supertrace.  Fix \(p\in M\).  Choose geodesic normal coordinates,
synchronous orthonormal frame, and radial gauges for \(S\) and \(E\).  In these
coordinates the Taylor expansions of the two connections begin with their
curvatures:
\begin{equation}
  \nabla_a^S
  =
  \partial_a+\frac14 R_{ab cd}(p)x^b\gamma^{cd}+O(|x|^2),
  \qquad
  \nabla_a^E
  =
  \partial_a+\frac12F^E_{ab}(p)x^b+O(|x|^2).
  \label{eq:index-radial-gauge-expansion}
\end{equation}
Rescale \(x=t^{1/2}y\).  Under Getzler scaling the differential operator
keeps the terms whose combined degree in \(t^{1/2}\), derivatives, and
Clifford variables can contribute to the top Clifford degree \(2n\); all
higher Taylor coefficients are suppressed by positive powers of \(t^{1/2}\)
in the diagonal heat kernel.  Lemma
\ref{lem:index-lichnerowicz} then gives the constant-curvature model
operator
\begin{equation}
  H_p
  =
  -\sum_a
  \left(
  \partial_{y^a}
  +\frac14 R_{ab}(p)y^b
  \right)^2
  +\frac12\gamma^{ab}F^E_{ab}(p),
  \label{eq:index-getzler-model-operator}
\end{equation}
where \(R_{ab}(p)\) is viewed as the skew endomorphism of \(T_pM\) with
matrix entries \(R_{ab cd}(p)\gamma^{cd}\) in the Clifford factor.  The
ordinary scalar-curvature term in \eqref{eq:index-lichnerowicz} has Clifford
degree \(0\), hence cannot by itself contribute to the supertrace in
dimension \(2n\); the missing gravitational contribution is instead produced
by the harmonic-oscillator determinant in \(H_p\).

The heat kernel of \eqref{eq:index-getzler-model-operator} is the Mehler
kernel for a constant magnetic oscillator, tensored with the finite
endomorphism exponential \(\exp(-\frac12\gamma^{ab}F^E_{ab})\).  Diagonalizing
the skew curvature into two-planes gives the pointwise supertrace
\begin{equation}
  \lim_{t\to0^+}{\rm str}\,K_t(p,p)\,d{\rm vol}_g(p)
  =
  \left[
  \det\nolimits^{1/2}
  \left(\frac{R/4\pi}{\sinh(R/4\pi)}\right)
  {\rm tr}_E
  \exp\!\left(\frac{\ii F_E}{2\pi}\right)
  \right]_{2n}.
  \label{eq:index-getzler-mehler-density}
\end{equation}
The determinant factor is the product over the curvature two-planes of the
Mehler ratio
\[
  \frac{u/2}{\sinh(u/2)},
\]
with \(u\) the corresponding Chern-Weil curvature variable.  The Clifford
supertrace extracts exactly the top exterior degree; lower Clifford degree
terms vanish because \({\rm str}_S\) is zero away from degree \(2n\).  Equation
\eqref{eq:index-getzler-mehler-density} is precisely
\(\left[\widehat A(TM)\operatorname{ch}(E)\right]_{2n}\).  Integrating over
\(M\) and using \eqref{eq:index-mckean-singer} proves
\eqref{eq:index-atiyah-singer}.
\end{proof}

This theorem is a theorem of differential geometry, not an axiom of QFT.  The
derivation above identifies which part is analytic heat-kernel control and
which part is local Clifford/Chern-Weil algebra.  The QFT consequences below
are then zero-mode selection rules, anomaly polynomials, determinant lines,
and inflow.

In four dimensions the formula becomes
\begin{equation}
  \operatorname{Ind}(D_A^+)
  =
  \int_M
  \left[
  \operatorname{ch}_2(E)
  -
  \frac{\operatorname{rank}(E)}{24}p_1(TM)
  \right].
  \label{eq:index-four-dimensional-formula}
\end{equation}
For an \(SU(N)\) bundle \(E_R\) associated to a representation \(R\),
\begin{equation}
  c_1(E_R)=0,
  \qquad
  \int_M\operatorname{ch}_2(E_R)=T_\Delta(R)\,k .
  \label{eq:index-trace-delta-instanton-number}
\end{equation}
Here \(k\in\mathbb Z\) is the instanton number in the fundamental bundle and
\(T_\Delta(R)\) is the quadratic index in the monograph's trace-delta
normalization
\[
  \operatorname{tr}_R(t^a_Rt^b_R)=T_\Delta(R)\delta^{ab},
  \qquad
  T_\Delta(\square)=1.
\]
Thus \(T_\Delta({\rm adj})=2N\).  This is the same normalization compatible
with the Yang-Mills kinetic term
\[
  \frac{1}{4g^2}\int\operatorname{tr}_\Delta(F_{\mu\nu}F^{\mu\nu}).
\]

\section{Four-Dimensional Examples}

On a spin four-manifold with \(p_1(TM)=0\), a unit \(SU(N)\) instanton gives
\[
  \operatorname{Ind}(D_{\square}^+)=1,
  \qquad
  \operatorname{Ind}(D_{\rm adj}^+)=2N .
\]
For \(SU(2)\), the fundamental Weyl fermion has one complex zero mode in the
orientation convention of \eqref{eq:index-trace-delta-instanton-number}, and
an adjoint Weyl fermion has four complex zero modes.  Reversing the
orientation of the gauge instanton reverses the sign of the index and swaps
the chirality of the zero modes.

For an Abelian line bundle \(L^q\) of charge \(q\),
\[
  \operatorname{ch}(L^q)=e^{q x},
  \qquad x=c_1(L).
\]
On \(T^4\) with \(p_1(TT^4)=0\), take integral two-forms \(a,b\) with
\[
  \int_{T^4}a\wedge b=1,
  \qquad a\wedge a=b\wedge b=0,
\]
and \(x=m_1a+m_2b\).  Then
\begin{equation}
  \operatorname{Ind}(D_{L^q}^+)
  =
  \int_{T^4}\frac12q^2x^2
  =
  q^2m_1m_2 .
  \label{eq:index-u1-t4-example}
\end{equation}
This example is useful because it is completely Abelian: the index is the
intersection pairing of magnetic fluxes through complementary two-tori.

\section{Zero Modes and Berezin Selection Rules}

Let \(\psi\) be a left-handed Weyl fermion valued in \(E\).  Decompose it in
an orthonormal eigenbasis of the coupled Dirac operator.  If
\(\{\zeta_1,\ldots,\zeta_m\}\) is a basis of \(\ker D_A^+\), then the
Grassmann field has zero-mode coordinates
\[
  \psi_0(x)=\sum_{i=1}^m\eta_i\zeta_i(x),
  \qquad \eta_i\eta_j=-\eta_j\eta_i .
\]

\paragraph{Finite-dimensional zero-mode rule.}
After nonzero modes have been paired, the zero-mode part of a fermionic path
integral is
\[
  \int d\eta_m\cdots d\eta_1\,\mathcal O(\eta_1,\ldots,\eta_m).
\]
It vanishes unless the insertion \(\mathcal O\) contains the top monomial
\(\eta_1\cdots\eta_m\).  Therefore a background with \(m\) unpaired Weyl
zero modes requires \(m\) corresponding fermion insertions for a nonzero
correlator.
Berezin integration is the linear functional fixed by
\[
  \int d\eta\,1=0,
  \qquad
  \int d\eta\,\eta=1.
\]
For \(m\) generators it extracts the coefficient of
\(\eta_1\cdots\eta_m\) in the declared order.  All lower-degree monomials are
annihilated by at least one of the integrations.

For a vectorlike Dirac fermion in representation \(R\), written in
two-component language as left-handed fields in \(R\) and \(\overline R\),
a charge-\(k\) instanton produces \(T_\Delta(R)k\) zero modes in each of the
two conjugate chiral variables, with chirality determined by the instanton
orientation.  The resulting finite-dimensional Berezin integral is the
precise origin of the fermion-number selection rule usually encoded by the
't Hooft vertex.

\section{Anomaly Polynomial and Descent}

For a chiral fermion in \(2n\) dimensions, the anomaly polynomial is the
\((2n+2)\)-form
\begin{equation}
  I_{2n+2}
  =
  \left[\widehat A(TM)\operatorname{ch}(E)\right]_{2n+2}.
  \label{eq:index-anomaly-polynomial}
\end{equation}
Choose a Chern-Simons form \(I_{2n+1}\) satisfying
\[
  dI_{2n+1}=I_{2n+2}.
\]
For an infinitesimal background gauge transformation with parameter
\(\alpha\),
\[
  \delta_\alpha I_{2n+1}=dI_{2n}^{(1)}(\alpha).
\]
The local anomalous variation of the Euclidean effective action is
\begin{equation}
  \delta_\alpha W[A,g]
  =
  2\pi\ii\int_M I_{2n}^{(1)}(\alpha),
  \label{eq:index-local-descent-variation}
\end{equation}
with an overall sign changed if the opposite chirality convention is chosen.
The sign is not independent data; it is fixed by the orientation of
\(D_A^+\), the definition of \(W=-\log Z\), and the chirality assignment of
the fermion.

In four-dimensional QFT, the relevant six-form is
\begin{equation}
  I_6
  =
  \operatorname{ch}_3(E)
  -
  \frac{p_1(TM)}{24}\operatorname{ch}_1(E).
  \label{eq:index-four-dimensional-anomaly-polynomial}
\end{equation}
The first term gives cubic gauge anomalies.  The second gives the mixed
gauge-gravitational anomaly for a \(U(1)\) background.  For a semisimple
group in a traceless representation, \(\operatorname{ch}_1(E)=0\), so the
mixed term is absent.  A background symmetry can be gauged only after the
total anomaly line has been trivialized; cancellation of the local polynomial
is necessary but does not by itself eliminate global anomalies.

\section{Determinant Line and Families}

As the metric and background connection vary, the chiral fermion determinant
is not naturally a complex number.  It is a section of the determinant line
\[
  \operatorname{Det}D
  =
  \bigwedge\nolimits^{\max}\ker D_A^+
  \otimes
  \left(\bigwedge\nolimits^{\max}\ker D_A^-\right)^\ast
\]
over background-field space.  The local anomaly is the curvature of this
line with its natural connection.  Global anomalies are its holonomies around
loops.  Thus a well-defined background-field partition function requires a
trivialization of the determinant line together with the vanishing of the
local curvature obstruction.
The source traces for the determinant-line and eta-holonomy formulation used
in the next chapter are Bismut--Freed and Dai--Freed
\cite{BismutFreed1986,DaiFreed1994}.

The next chapter develops the eta-invariant formula for these holonomies.
The present chapter supplies the local and zero-mode input: the index counts
the imbalance of chiral zero modes, while the family of such indices
organizes the determinant line.

\begin{openproblem}[Index-theoretic anomaly line in interacting QFT]
\label{op:index-anomaly-line-interacting-qft}
Extend the determinant-line and index-theoretic description from free chiral
fermions coupled to backgrounds to interacting continuum QFTs with composite
currents, contact terms, and extended operators, while retaining an explicit
comparison with local anomaly descent and global anomaly holonomy.
\end{openproblem}
