\chapter{Semiclassical Backreaction And Stress-Tensor Fluctuations}
\label{ch:curved-semiclassical-backreaction}

\ConventionsLorentzian

QFT on a fixed curved background supplies local observables, states, and a
renormalized stress tensor.  Once the stress tensor is used as a source for
geometry, the metric is no longer a passive parameter.  A semiclassical
problem must therefore specify both a locally covariant quantum field theory
on each admissible metric and a gravitational effective action whose metric
variation is equated to the stress-tensor expectation value.  This chapter
develops that statement, the associated linear response theory, and the
stress-tensor fluctuation data needed for stochastic backreaction.
Its upstream inputs are explicit: a microlocal Hadamard state class, a
locally covariant algebra, a pAQFT or free-field prescription for composite
stress tensors and time-ordered products, and the point-splitting/anomaly
finite-counterterm choices.  Backreaction begins only after those choices are
joined to gravitational EFT coordinates and a response/noise window.

The physical question is not simply whether the formal equation
\(E^{\rm grav}=\langle T^{\rm ren}\rangle\) can be written.  It is whether a
specified quantum state, stress tensor, and stress-tensor fluctuation law
produce a controlled metric observable.  The finite chain used in this chapter
is
\begin{equation}
  (\mathcal A_g,\omega,T_{\mathcal S}^{\rm ren})
  \longrightarrow
  (E_\theta^{\rm grav},\mathsf T_{g\to g+h})
  \longrightarrow
  (j,N,K^{\rm ret})
  \longrightarrow
  (h_*,\mathsf C_h)
  \longrightarrow
  X[g_0+h].
  \label{eq:semiclassical-observable-chain}
\end{equation}
The entries mean, respectively, the locally covariant algebra/state/stress
tensor in a fixed renormalization scheme, the gravitational EFT coordinates
and state-transport rule, the Ward-clean source/noise/retarded response data,
the mean metric shift and Einstein--Langevin covariance, and the retained
metric observable.  A formal mean equation, a small coordinate perturbation, or
a projected noise covariance is therefore not yet a physical semiclassical
prediction; the observable and its uncertainty are part of the output.

\section{Metric Variation And Curvature Convention}

The curvature convention in this chapter is
\begin{equation}
  [\nabla_\mu,\nabla_\nu]X^\rho
  =
  R^\rho{}_{\sigma\mu\nu}X^\sigma,
  \qquad
  R_{\mu\nu}=R^\rho{}_{\mu\rho\nu},
  \qquad
  G_{\mu\nu}=R_{\mu\nu}-\frac12Rg_{\mu\nu}.
  \label{eq:semiclassical-curvature-convention}
\end{equation}
Metric variations are taken with respect to \(g^{\mu\nu}\).  For a
renormalized source functional \(W[g]\), define the stress tensor by
\begin{equation}
  \delta W[g]
  =
  -\frac12
  \int_M \dd\operatorname{vol}_g\,
  \langle T_{\mu\nu}^{\rm ren}\rangle_g\,\delta g^{\mu\nu}.
  \label{eq:semiclassical-stress-source-definition}
\end{equation}
The minus sign is the same sign that gives
\(\delta S_{\rm matter}=-(1/2)\int T_{\mu\nu}\delta g^{\mu\nu}\) in the
classical limit with mostly-plus metric.

\paragraph{Ward identity from diffeomorphism covariance.}
Suppose \(W[g]\) is invariant under compactly supported diffeomorphisms and
has metric variation of the form
\eqref{eq:semiclassical-stress-source-definition}.  Then
\begin{equation}
  \nabla^\mu\langle T_{\mu\nu}^{\rm ren}\rangle_g=0
  \label{eq:semiclassical-stress-conservation}
\end{equation}
as a distribution.

Let \(X^\mu\) be a compactly supported vector field and let
\(\psi_s\) be its flow.  Diffeomorphism invariance gives
\[
  0
  =
  \left.\frac{\dd}{\dd s}W[\psi_s^*g]\right|_{s=0}.
\]
Since
\[
  \delta_X g_{\mu\nu}=\nabla_\mu X_\nu+\nabla_\nu X_\mu,
  \qquad
  \delta_X g^{\mu\nu}=-(\nabla^\mu X^\nu+\nabla^\nu X^\mu),
\]
equation \eqref{eq:semiclassical-stress-source-definition} gives
\[
  0
  =
  \int_M \dd\operatorname{vol}_g\,
  \langle T_{\mu\nu}^{\rm ren}\rangle_g\nabla^\mu X^\nu .
\]
Integrating by parts and using compact support of \(X\) gives
\[
  0
  =
  -\int_M \dd\operatorname{vol}_g\,
  X^\nu\nabla^\mu\langle T_{\mu\nu}^{\rm ren}\rangle_g .
\]
Since \(X\) is arbitrary, the distribution in
\eqref{eq:semiclassical-stress-conservation} vanishes.

If the theory has a diffeomorphism anomaly, the right side of
\eqref{eq:semiclassical-stress-conservation} is the local anomaly.  In the
ordinary scalar, spinor, and vector examples considered in the fixed
background chapters, diffeomorphism covariance is imposed on the
renormalized stress tensor; trace anomalies are compatible with
\eqref{eq:semiclassical-stress-conservation}.

\section{Renormalized Stress Tensor}

Let \((M,g)\) be globally hyperbolic and let \(\omega\) be a Hadamard state
of a locally covariant QFT.  The renormalized stress tensor is an
operator-valued distribution
\[
  T_{\mu\nu}^{\rm ren}(f^{\mu\nu}),
  \qquad
  f^{\mu\nu}\in C_c^\infty(M,S^2TM),
\]
or, in perturbative algebraic language, a local covariant Wick polynomial
valued in the algebra constructed from the Hadamard parametrix.  Its
expectation value is
\[
  \langle T_{\mu\nu}^{\rm ren}(x)\rangle_{\omega,g}.
\]
For Hadamard states this is a smooth tensor field after point splitting and
after the finite local curvature terms have been fixed.  For a scalar field,
the subtraction is the one described in
Chapter~\ref{ch:point-splitting-stress-tensor-renormalization}; for an
interacting pAQFT model it is the local Wick-polynomial and
time-ordered-product renormalization of
Chapter~\ref{ch:curved-perturbative-algebraic-qft}, on the local/equicausal
functional domain where the Peierls bracket, star product, and time-slice
maps are closed.  A bare microcausal wavefront check is not enough data for
an interacting stress tensor to serve as a conserved backreaction source.

The finite ambiguity in four spacetime dimensions has the form
\begin{equation}
  \langle T_{\mu\nu}^{\rm ren}\rangle
  \longmapsto
  \langle T_{\mu\nu}^{\rm ren}\rangle
  +
  c_0m^4g_{\mu\nu}
  +
  c_1m^2G_{\mu\nu}
  +
  c_2H^{(1)}_{\mu\nu}
  +
  c_3H^{(2)}_{\mu\nu},
  \label{eq:semiclassical-stress-finite-ambiguity}
\end{equation}
for a massive scalar model, with the evident modifications when additional
background couplings are present.  The tensors \(H^{(i)}_{\mu\nu}\) are
local conserved curvature tensors defined in the next section.  Thus
renormalizing the stress tensor and choosing gravitational EFT coordinates
are not independent operations.

\section{Gravitational Effective Action}

In four dimensions, take the local gravitational part of the renormalized
effective action to be
\begin{equation}
  \Gamma_{\rm grav}[g]
  =
  \int_M\dd\operatorname{vol}_g
  \left[
    \frac{1}{16\pi G_N}(R-2\Lambda)
    +
    \frac{\alpha}{2}R^2
    +
    \frac{\beta}{2}R_{\rho\sigma}R^{\rho\sigma}
  \right],
  \label{eq:semiclassical-gravitational-action}
\end{equation}
up to boundary terms and topological combinations.  The factor \(1/2\) in
the curvature-squared terms fixes the normalization
\begin{equation}
  E_{\mu\nu}^{\rm grav}
  :=
  \frac{2}{\sqrt{|g|}}\frac{\delta\Gamma_{\rm grav}}{\delta g^{\mu\nu}}
  =
  \frac1{8\pi G_N}(G_{\mu\nu}+\Lambda g_{\mu\nu})
  +
  \alpha H^{(1)}_{\mu\nu}
  +
  \beta H^{(2)}_{\mu\nu}.
  \label{eq:semiclassical-gravitational-euler-tensor}
\end{equation}
\paragraph{Curvature-squared variational formulas.}
With the curvature convention
\eqref{eq:semiclassical-curvature-convention}, and with the same
metric-variation normalization used in
Chapter~\ref{ch:point-splitting-stress-tensor-renormalization}, the
curvature-squared Euler tensors in
\eqref{eq:semiclassical-gravitational-euler-tensor} are
\begin{equation}
\begin{aligned}
  H^{(1)}_{\mu\nu}
  &=
  2RR_{\mu\nu}
  -\frac12g_{\mu\nu}R^2
  +
  2g_{\mu\nu}\Box R
  -
  2\nabla_\mu\nabla_\nu R,\\
  H^{(2)}_{\mu\nu}
  &=
  2R_{\mu\rho\nu\sigma}R^{\rho\sigma}
  -\frac12g_{\mu\nu}R_{\rho\sigma}R^{\rho\sigma}
  -
  \nabla_\mu\nabla_\nu R
  +
  \Box R_{\mu\nu}
  +
  \frac12g_{\mu\nu}\Box R .
\end{aligned}
  \label{eq:semiclassical-H-tensors}
\end{equation}
Here \(\Box=g^{\rho\sigma}\nabla_\rho\nabla_\sigma\).

These formulae are variational identities fixing the local
gravitational-coordinate normalization, not an additional structural theorem.
The basic first variations are
\begin{equation}
  \delta\sqrt{|g|}
  =
  -\frac12\sqrt{|g|}\,g_{\mu\nu}\delta g^{\mu\nu},
  \qquad
  \delta R
  =
  R_{\mu\nu}\delta g^{\mu\nu}
  +
  g_{\mu\nu}\Box\delta g^{\mu\nu}
  -
  \nabla_\mu\nabla_\nu\delta g^{\mu\nu},
  \label{eq:semiclassical-basic-metric-variations}
\end{equation}
after discarding boundary terms.  Applying these identities to
\(\frac12\int\sqrt{|g|}R^2\) gives the first line of
\eqref{eq:semiclassical-H-tensors}.

For \(\frac12\int\sqrt{|g|}R_{\rho\sigma}R^{\rho\sigma}\), use
\begin{equation}
  \delta R_{\mu\nu}
  =
  \frac12\left(
    -\Box\delta g_{\mu\nu}
    -\nabla_\mu\nabla_\nu(g^{\rho\sigma}\delta g_{\rho\sigma})
    +
    \nabla^\rho\nabla_\mu\delta g_{\rho\nu}
    +
    \nabla^\rho\nabla_\nu\delta g_{\rho\mu}
  \right),
  \label{eq:semiclassical-ricci-variation}
\end{equation}
convert \(\delta g_{\mu\nu}\) to
\(-g_{\mu\rho}g_{\nu\sigma}\delta g^{\rho\sigma}\), integrate derivatives
off \(\delta g^{\mu\nu}\), and commute derivatives on \(R_{\mu\nu}\).  The
minus sign in this conversion is what fixes the derivative signs.  In the
half-normalized action the integrated Palatini term contributes
\[
  -\nabla_\mu\nabla_\nu R+\Box R_{\mu\nu}
  +\frac12g_{\mu\nu}\Box R
\]
to \(H^{(2)}_{\mu\nu}\).  The commutator terms combine with the explicit
metric variation of \(R^{\rho\sigma}\) into
\(2R_{\mu\rho\nu\sigma}R^{\rho\sigma}\), giving the second line of
\eqref{eq:semiclassical-H-tensors}.  The opposite derivative signs would be
the variation of neither the displayed half-normalized action nor the
full \(\int\sqrt{|g|}R_{\rho\sigma}R^{\rho\sigma}\) action in the
\(g^{\mu\nu}\)-variation convention.

The tensors \(H^{(1)}\) and \(H^{(2)}\) are conserved.  This follows without
component calculation: each is the Euler tensor of a diffeomorphism-invariant
local functional, so the Ward-identity calculation above applies with the
functional replacing \(W\).  Their traces in \(D=4\) are
\begin{equation}
  g^{\mu\nu}H^{(1)}_{\mu\nu}=6\Box R,
  \qquad
  g^{\mu\nu}H^{(2)}_{\mu\nu}=2\Box R,
  \label{eq:semiclassical-H-traces}
\end{equation}
using the convention in \eqref{eq:semiclassical-H-tensors}.  These traces
are total derivatives, as required by the constant-Weyl-scale invariance of
the curvature-squared actions in four dimensions.  For the un-halved Ricci
squared action, the Weyl variation
\(\delta g^{\mu\nu}=-2\sigma g^{\mu\nu}\) gives
\[
  \delta_\sigma
  \int\sqrt{|g|}\,R_{\rho\sigma}R^{\rho\sigma}
  =
  -4\int\sqrt{|g|}\,\sigma\Box R
\]
after integrating by parts, which is the same sign check.

\section{Semiclassical Einstein Equation}

In the control-level classification of
Section~\ref{sec:curved-qft-control-levels}, the equation below is a
conditional semiclassical statement, not a theorem of interacting quantum
gravity.  It becomes a physical backreaction claim only as the first stage of
\eqref{eq:semiclassical-observable-chain}: the locally covariant algebra,
admissible state class, renormalized stress tensor, gravitational EFT
coordinates, state-transport rule, and response window must be fixed together.

A semiclassical solution is not only a metric.  It is a pair \((g,\omega)\)
such that \(\omega\) is an admissible Hadamard state of the QFT constructed
on \((M,g)\) and
\begin{equation}
  E_{\mu\nu}^{\rm grav}[g]
  =
  \langle T_{\mu\nu}^{\rm ren}\rangle_{\omega,g}.
  \label{eq:semiclassical-einstein-equation-expanded}
\end{equation}
Written out,
\begin{equation}
  \frac1{8\pi G_N}(G_{\mu\nu}+\Lambda g_{\mu\nu})
  +
  \alpha H^{(1)}_{\mu\nu}
  +
  \beta H^{(2)}_{\mu\nu}
  =
  \langle T_{\mu\nu}^{\rm ren}\rangle_{\omega,g}.
  \label{eq:semiclassical-einstein-equation}
\end{equation}
The left side is conserved by diffeomorphism invariance.  Therefore the
right side must be the conserved stress tensor of the same renormalization
scheme.  A change of stress-tensor scheme of the form
\eqref{eq:semiclassical-stress-finite-ambiguity} is equivalently a finite
change of \(G_N,\Lambda,\alpha,\beta\).  It is not a physical change until a
renormalization condition has fixed these coordinates.

\begin{proposition}[Finite scheme transport of the semiclassical equation]
\label{prop:semiclassical-finite-scheme-transport}
Let \(\mathcal S\) be a stress-tensor prescription and write
\[
  \mathcal R_{\mathcal S,\theta}(g,\omega)
  =
  E^{\rm grav}_\theta[g]
  -
  \langle T_{\mu\nu}^{\mathcal S}\rangle_{\omega,g}
\]
for the semiclassical residual, where
\[
  E^{\rm grav}_\theta
  =
  M_{\rm P}^2G_{\mu\nu}
  +
  \lambda_{\rm grav}g_{\mu\nu}
  +
  \alpha H^{(1)}_{\mu\nu}
  +
  \beta H^{(2)}_{\mu\nu},
  \qquad
  M_{\rm P}^2=(8\pi G_N)^{-1},
  \quad
  \lambda_{\rm grav}=M_{\rm P}^2\Lambda .
\]
Suppose \(\mathcal S'\) differs from \(\mathcal S\) by the finite local
curvature tensor
\begin{equation}
  A_{\mu\nu}[g]
  =
  a_0m^4g_{\mu\nu}
  +
  a_1m^2G_{\mu\nu}
  +
  a_2H^{(1)}_{\mu\nu}
  +
  a_3H^{(2)}_{\mu\nu},
  \qquad
  \langle T_{\mu\nu}^{\mathcal S'}\rangle
  =
  \langle T_{\mu\nu}^{\mathcal S}\rangle
  +
  A_{\mu\nu}.
  \label{eq:semiclassical-finite-scheme-shift-A}
\end{equation}
Then the same semiclassical equation is represented in the primed scheme by
\begin{equation}
\begin{aligned}
  M_{\rm P}^{\prime\,2}
  &=
  M_{\rm P}^{2}+a_1m^2,
  &
  \lambda_{\rm grav}'
  &=
  \lambda_{\rm grav}+a_0m^4,
  \\
  \alpha'&=\alpha+a_2,
  &
  \beta'&=\beta+a_3 ,
\end{aligned}
  \label{eq:semiclassical-finite-gravity-coordinate-transport}
\end{equation}
or, equivalently,
\begin{equation}
  E^{\rm grav}_{\theta'}[g]
  =
  E^{\rm grav}_{\theta}[g]+A_{\mu\nu}[g],
  \qquad
  \mathcal R_{\mathcal S',\theta'}(g,\omega)
  =
  \mathcal R_{\mathcal S,\theta}(g,\omega).
  \label{eq:semiclassical-residual-scheme-invariance}
\end{equation}
\end{proposition}

\begin{proof}
The four tensors in
\eqref{eq:semiclassical-finite-scheme-shift-A} are exactly the local conserved
basis appearing in
Proposition~\ref{prop:finite-local-freedom-scalar-stress-tensor}, with the
normalization of
\eqref{eq:semiclassical-gravitational-euler-tensor}.  The coordinate
replacement
\eqref{eq:semiclassical-finite-gravity-coordinate-transport} adds the same
tensor \(A_{\mu\nu}\) to the gravitational Euler tensor.  Substituting
\(E_\theta+A\) and \(T_{\mathcal S}+A\) into
\(\mathcal R=E-T\) gives
\eqref{eq:semiclassical-residual-scheme-invariance}.
\end{proof}

The proposition is a coordinate statement, not a license to move only one side
of the equation.  If
\(\langle T_{\mu\nu}\rangle\) is shifted by \(A_{\mu\nu}\) while
\((M_{\rm P}^2,\lambda_{\rm grav},\alpha,\beta)\) are held fixed, the
semiclassical residual changes by \(-A_{\mu\nu}\).  If the gravitational
coordinates are shifted without changing the stress-tensor prescription, the
residual changes by \(+A_{\mu\nu}\).  The same bookkeeping is required after
linearization:
\begin{equation}
  \delta E_{\theta'}^{\rm grav}[h]
  -
  \delta\langle T^{\mathcal S'}\rangle[h]
  =
  \delta E_{\theta}^{\rm grav}[h]
  -
  \delta\langle T^{\mathcal S}\rangle[h],
  \qquad
  \delta\langle T^{\mathcal S'}\rangle[h]
  =
  \delta\langle T^{\mathcal S}\rangle[h]
  +
  \delta A[h].
  \label{eq:semiclassical-linear-scheme-transport}
\end{equation}
Thus the local contact part of the stress response and the local
curvature-squared part of the gravitational operator must be transported
together.  A c-number curvature shift has no separated commutator and does not
change connected stress fluctuations:
\begin{equation}
  N_{\mathcal S'}(f,k)
  =
  N_{\mathcal S}(f,k)
  \label{eq:semiclassical-cnumber-noise-scheme-invariance}
\end{equation}
for the deterministic shift
\eqref{eq:semiclassical-finite-scheme-shift-A}.  This statement does not apply
to finite mixing with nontrivial composite operators; those changes modify the
connected noise by the cross terms described in
\eqref{eq:semiclassical-noise-finite-renormalization}.  The physical lesson is
that the mean equation, the local linear response, and the connected
Einstein--Langevin noise are different derivatives of one renormalized
stress-tensor coordinate system.  Mixing coordinates across them is a change of
the model, not a harmless relabeling.

\begin{definition}[Backreaction datum]
\label{def:semiclassical-backreaction-datum}
A fixed-order semiclassical backreaction datum consists of
\[
  \mathfrak B
  =
  (M,\mathcal C,\mathcal A_g,\mathcal S_g,
    T_{\mu\nu}^{\rm ren},G_N,\Lambda,\alpha,\beta,\mathcal I).
\]
Here \(\mathcal C\) is the class of metrics considered,
\(\mathcal A_g\) is the locally covariant QFT algebra assigned to \(g\),
\(\mathcal S_g\) is the admissible state space, \(T^{\rm ren}\) is the
renormalized stress tensor in a fixed scheme, and \(\mathcal I\) is the
initial or boundary data for the metric and state.
\end{definition}

The datum matters because the equation is state dependent.  For example, the
same metric may solve \eqref{eq:semiclassical-einstein-equation} for one
state and fail for another.  Conversely, changing the metric changes the
algebra and the Hadamard condition, so the state must be transported or
reconstructed as part of the problem.

\section{Linear Response}

\begin{proposition}[Retarded stress-tensor response at separated points]
\label{prop:semiclassical-retarded-stress-linear-response}
Let the locally covariant theory be fixed on a globally hyperbolic background
\((M,g)\), let \(\omega\) be an admissible Hadamard state, and let the
renormalized stress tensor and its time-ordered square be chosen in one
renormalization scheme.  Vary the inverse metric by a compactly supported
smooth tensor \(h^{\mu\nu}=\delta g^{\mu\nu}\).  The metric perturbation
couples to the stress tensor through
\begin{equation}
  \delta S_{\rm source}
  =
  -\frac12
  \int_M\dd\operatorname{vol}_g\,
  h^{\rho\sigma}T_{\rho\sigma}^{\rm ren}.
  \label{eq:semiclassical-metric-source-coupling}
\end{equation}
The first variation of the stress-tensor expectation has the form
\begin{equation}
  \delta\langle T_{\mu\nu}(x)\rangle
  =
  \frac12\int_M\dd\operatorname{vol}_g(y)\,
  G^{\rm ret}_{\mu\nu,\rho\sigma}(x,y)h^{\rho\sigma}(y)
  +
  \delta T_{\mu\nu}^{\rm loc}(x).
  \label{eq:semiclassical-linear-response}
\end{equation}
The separated-point retarded kernel is
\begin{equation}
  G^{\rm ret}_{\mu\nu,\rho\sigma}(x,y)
  =
  \ii\,\Theta(x\succ y)\,
  \omega\!\left(
    [T_{\mu\nu}^{\rm ren}(x),T_{\rho\sigma}^{\rm ren}(y)]
  \right),
  \label{eq:semiclassical-retarded-stress-kernel}
\end{equation}
where \(x\succ y\) means \(x\in J^+(y)\).  The local term
\(\delta T_{\mu\nu}^{\rm loc}\) contains variations of the point-splitting
subtraction, curvature counterterms, and contact terms in the definition of
the time-ordered stress-tensor product.
\end{proposition}

\begin{proof}[Derivation of the retarded commutator]
The clean statement is first made for a smeared observable
\(A(f)=\int A(x)f(x)\) whose support is disjoint from the support of the
local counterterm extension of the stress-tensor product.  In the
interaction-picture or pAQFT notation, the relative \(S\)-matrix for the
source \(V=\delta S_{\rm source}\) gives, to first order,
\[
  \delta\omega(A(f))
  =
  \ii\,\omega\!\left([V_{J^-(\operatorname{supp}f)},A(f)]\right)
  +
  \delta_{\rm loc}\omega(A(f)).
\]
Only the part of \(V\) in the causal past of \(\operatorname{supp}f\)
appears; this is the retarded prescription.  The local term
\(\delta_{\rm loc}\) records two effects that cannot be read from separated
points: the explicit variation of the composite field \(A\) with the metric
and the chosen extension of time-ordered products to the diagonal.

Set \(A=T_{\mu\nu}^{\rm ren}\) and insert
\eqref{eq:semiclassical-metric-source-coupling}.  For \(y\in J^-(x)\),
\[
  \ii\left[
    -\frac12 h^{\rho\sigma}(y)T_{\rho\sigma}^{\rm ren}(y),
    T_{\mu\nu}^{\rm ren}(x)
  \right]
  =
  \frac12 h^{\rho\sigma}(y)\,
  \ii\,[T_{\mu\nu}^{\rm ren}(x),T_{\rho\sigma}^{\rm ren}(y)] .
\]
After unsmearing away from the diagonal, this gives
\eqref{eq:semiclassical-linear-response} with the kernel
\eqref{eq:semiclassical-retarded-stress-kernel}.  Microcausality makes the
commutator vanish for spacelike separated \(x,y\); the restriction
\(\Theta(x\succ y)\) removes the advanced part and is therefore the unique
causal linear response compatible with the sign convention
\(\delta W=-(1/2)\int T_{\rho\sigma}\delta g^{\rho\sigma}\).  The remaining
diagonal ambiguity is exactly the local term
\(\delta T_{\mu\nu}^{\rm loc}\).
\end{proof}

\paragraph{Linearized diffeomorphism Ward identity.}
The response \eqref{eq:semiclassical-linear-response} is not, by itself, an
ordinary transverse two-tensor distribution.  Let
\[
  \bar T_{\mu\nu}:=\langle T_{\mu\nu}^{\rm ren}\rangle_g,
  \qquad
  \mathcal R_{\mu\nu}[h]
  :=
  \frac12\int_M\dd\operatorname{vol}_g(y)\,
  G^{\rm ret}_{\mu\nu,\rho\sigma}(x,y)h^{\rho\sigma}(y)
  +
  \delta T_{\mu\nu}^{\rm loc}[h]
\]
denote the metric derivative of the stress expectation in the fixed
renormalization and state-transport prescription.  Varying
\eqref{eq:semiclassical-stress-conservation} gives
\begin{equation}
  \nabla^\mu\mathcal R_{\mu\nu}[h]
  +
  \mathcal C_\nu[h;\bar T]
  =
  0,
  \label{eq:semiclassical-linearized-stress-ward}
\end{equation}
where the one-point connection term is
\begin{equation}
\begin{aligned}
  \mathcal C_\nu[h;\bar T]
  &:=
  h^{\mu\alpha}\nabla_\alpha\bar T_{\mu\nu}
  -
  g^{\mu\alpha}
  (\delta_h\Gamma)^\lambda_{\alpha\mu}\bar T_{\lambda\nu}
  -
  g^{\mu\alpha}
  (\delta_h\Gamma)^\lambda_{\alpha\nu}\bar T_{\mu\lambda},
  \\
  (\delta_h\Gamma)^\lambda_{\alpha\beta}
  &:=
  \frac12 g^{\lambda\kappa}
  \left(
    \nabla_\alpha\gamma_{\beta\kappa}
    +
    \nabla_\beta\gamma_{\alpha\kappa}
    -
    \nabla_\kappa\gamma_{\alpha\beta}
  \right),
  \qquad
  \gamma_{\mu\nu}:=\delta g_{\mu\nu}
  =
  -g_{\mu\rho}g_{\nu\sigma}h^{\rho\sigma}.
\end{aligned}
  \label{eq:semiclassical-linearized-connection-contact}
\end{equation}
Thus the separated commutator in
\eqref{eq:semiclassical-retarded-stress-kernel} need not obey
\(\nabla^\mu G^{\rm ret}_{\mu\nu,\rho\sigma}=0\) as a distribution.  When the
divergence hits the retarded step function, equal-time stress commutators and
the extension of the time-ordered product to the diagonal produce precisely
the contact terms included in \(\delta T^{\rm loc}\), while
\(\mathcal C_\nu[h;\bar T]\) records the variation of the connection acting on
the nonzero background one-point stress tensor.

For a pure diffeomorphism perturbation, with the inverse-metric convention
\(h^{\mu\nu}=\mathcal L_Xg^{\mu\nu}\), covariance says
\begin{equation}
  \mathcal R_{\mu\nu}[\mathcal L_X g]
  =
  \mathcal L_X\bar T_{\mu\nu},
  \label{eq:semiclassical-pure-gauge-stress-response}
\end{equation}
not that the stress response vanishes.  The null direction belongs to the full
semiclassical residual.  On a background satisfying
\(E_{\mu\nu}^{\rm grav}[g]=\bar T_{\mu\nu}\), the complete linearized operator
\[
  \mathcal D_{\rm sc}h
  :=
  \delta E_{\mu\nu}^{\rm grav}[h]-\mathcal R_{\mu\nu}[h]
\]
obeys
\begin{equation}
  \mathcal D_{\rm sc}(\mathcal L_X g)
  =
  \mathcal L_X(E^{\rm grav}-\bar T)
  =
  0,
  \label{eq:semiclassical-full-hessian-gauge-identity}
\end{equation}
modulo the chosen gauge fixing and constraint projection.  This is the
diffeomorphism Ward identity relevant for backreaction: gauge degeneracy is a
property of the combined gravitational-plus-matter Hessian, not of the
separated retarded stress commutator.

The linearized semiclassical equation is therefore
\begin{equation}
  \delta E_{\mu\nu}^{\rm grav}[h]
  -
  \frac12\int_M\dd\operatorname{vol}_g(y)\,
  G^{\rm ret}_{\mu\nu,\rho\sigma}(x,y)h^{\rho\sigma}(y)
  =
  \delta\langle T_{\mu\nu}\rangle_{\rm state}
  +
  \delta T_{\mu\nu}^{\rm loc}.
  \label{eq:semiclassical-linearized-equation}
\end{equation}
The term \(\delta\langle T\rangle_{\rm state}\) denotes an independent
change of state at fixed metric; it is absent if the state is transported by
the chosen metric perturbation prescription.

\section{Noise Kernel}

Stress-tensor fluctuations are encoded by the unordered symmetrized connected
two-point distribution of the already renormalized stress tensor:
\begin{equation}
  N_{\mu\nu,\rho\sigma}(x,y)
  =
  \frac12\,
  \omega\!\left(
    \{T_{\mu\nu}^{\rm ren}(x),T_{\rho\sigma}^{\rm ren}(y)\}
  \right)_{\rm conn}.
  \label{eq:semiclassical-noise-kernel}
\end{equation}
More invariantly, fix the algebra product in which the smeared stress tensor
is an observable and define, for real compactly supported test tensors
\(f,k\),
\begin{equation}
  N(f,k)
  =
  \frac12\,\omega\!\left(
    \Delta T(f)\Delta T(k)+\Delta T(k)\Delta T(f)
  \right),
  \qquad
  \Delta T(f)=T(f)-\omega(T(f))\mathbf 1 .
  \label{eq:semiclassical-noise-bilinear-form}
\end{equation}
Equivalently, set
\[
  T(f)=\int_M\dd\operatorname{vol}_g\,f^{\mu\nu}T_{\mu\nu}^{\rm ren}.
\]
Then
\begin{equation}
  N(f,f)
  =
  \frac12\omega(\{T(f)-\omega(T(f)),T(f)-\omega(T(f))\})
  =
  \omega\!\left((T(f)-\omega(T(f)))^2\right)
  \ge0 .
  \label{eq:semiclassical-noise-positivity}
\end{equation}
Thus \(N\) is a positive semidefinite distributional covariance kernel on
the real test-tensor space where the unordered product of smeared stress
observables is defined.  If the bilinear form
\eqref{eq:semiclassical-noise-bilinear-form} is continuous on the chosen
test-tensor topology, the Schwartz kernel theorem gives a unique
bidistribution \(N_{\mu\nu,\rho\sigma}\).  Its wavefront and continuity domain
are hypotheses or consequences of that algebra/state construction; they are
not obtained by choosing an independent time-ordered extension at
coincidence.

Conservation is the distributional Ward identity
\begin{equation}
  \nabla_x^\mu N_{\mu\nu,\rho\sigma}(x,y)=0,
  \qquad
  \nabla_y^\rho N_{\mu\nu,\rho\sigma}(x,y)=0,
  \label{eq:semiclassical-noise-conservation}
\end{equation}
Away from the diagonal it follows immediately from
\(\nabla^\mu T_{\mu\nu}^{\rm ren}=0\).  On the diagonal it is a property of
the same unordered algebra product and state used in
\eqref{eq:semiclassical-noise-bilinear-form}.  One must not add an
independently chosen local term to \(N\) merely because a time-ordered product
has Epstein--Glaser extension freedom: such a term changes variances and can
destroy positivity.  C-number curvature ambiguities in
\(T_{\mu\nu}^{\rm ren}\) cancel from \(\Delta T\); genuine finite operator
mixing changes the stress observable itself and therefore changes \(N\) only
through the covariance, cross-covariance, and Ward checks of that new
observable.  The local metric-variation and time-ordered contact terms belong
to the retarded response kernel and the gravitational effective action, not to
an adjustable noise covariance.

\section{Stationary States And Fluctuation--Dissipation}

If \(g\) has a complete timelike Killing field \(K\) and \(\omega\) is a KMS
state at inverse temperature \(\beta\) for the associated time evolution,
the retarded kernel and the noise kernel are related.  For two smeared
stress-tensor observables \(A\) and \(B\), define
\[
  C_{AB}(t)=\omega(A(t)B),
  \qquad
  \rho_{AB}(t)=\omega([A(t),B]).
\]
Let tildes denote Fourier transforms in \(t\).  The KMS condition gives
\begin{equation}
  \widetilde C_{BA}(-\omega)
  =
  e^{-\beta\omega}\widetilde C_{AB}(\omega).
  \label{eq:semiclassical-kms-spectral-relation}
\end{equation}
Therefore the symmetrized spectrum and the commutator spectrum obey
\begin{equation}
  \widetilde N_{AB}(\omega)
  =
  \frac12\coth\!\left(\frac{\beta\omega}{2}\right)
  \widetilde\rho_{AB}(\omega),
  \label{eq:semiclassical-fdt}
\end{equation}
where
\[
  \widetilde N_{AB}(\omega)
  =
  \frac12\left(\widetilde C_{AB}(\omega)
  +\widetilde C_{BA}(-\omega)\right),
  \qquad
  \widetilde\rho_{AB}(\omega)
  =
  \widetilde C_{AB}(\omega)
  -\widetilde C_{BA}(-\omega).
\]
This follows directly from KMS analyticity and the definitions of the retarded
and symmetrized kernels.

\section{Einstein--Langevin Equation}

A stochastic semiclassical approximation introduces a real Gaussian
generalized random tensor \(\xi_{\mu\nu}\) with
\begin{equation}
  \mathbb E[\xi_{\mu\nu}(f^{\mu\nu})]=0,
  \qquad
  \mathbb E[\xi(f)\xi(k)]=N(f,k).
  \label{eq:semiclassical-random-source-definition}
\end{equation}
The existence of this Gaussian generalized random field follows from the
positivity \eqref{eq:semiclassical-noise-positivity}: \(N\) is a covariance
form on a nuclear test-function space, so the Bochner--Minlos theorem
constructs the corresponding Gaussian measure on distributions.

After gauge fixing the linearized gravitational operator, the
Einstein--Langevin equation is
\begin{equation}
  \mathcal L_{\mu\nu}{}^{\rho\sigma}h_{\rho\sigma}
  -
  \frac12\int_M\dd\operatorname{vol}_g(y)\,
  G^{\rm ret}_{\mu\nu,\rho\sigma}(x,y)h^{\rho\sigma}(y)
  =
  \xi_{\mu\nu}(x),
  \label{eq:semiclassical-einstein-langevin}
\end{equation}
where \(\mathcal L\) is the gauge-fixed variation of
\(E_{\mu\nu}^{\rm grav}\) plus the local stress-tensor response.  If
\(\mathcal G_{\rm ret}\) is a retarded inverse for the full linear operator,
then the induced metric covariance is
\begin{equation}
  \mathbb E[h_{\mu\nu}(x)h_{\rho\sigma}(y)]
  =
  \int
  \mathcal G_{{\rm ret}\,\mu\nu}{}^{\alpha\beta}(x,u)
  N_{\alpha\beta,\gamma\delta}(u,v)
  \mathcal G_{{\rm ret}\,\rho\sigma}{}^{\gamma\delta}(y,v).
  \label{eq:semiclassical-metric-covariance}
\end{equation}
This equation is meaningful only after the state, stress-tensor
renormalization, contact terms, gauge condition, and stochastic test space
have all been specified.

\paragraph{Pushforward covariance of the Einstein--Langevin source.}
Let \(\mathcal E\) be the real nuclear test-tensor space on which the noise
covariance \(N\) is defined, and let \(\mathcal E'\) be its distributional
dual.  Suppose the gauge-fixed linear operator in
\eqref{eq:semiclassical-einstein-langevin} has a continuous retarded
solution map
\begin{equation}
  \mathcal G_{\rm ret}:\mathcal E'\longrightarrow \mathcal H ,
  \label{eq:semiclassical-retarded-solution-map}
\end{equation}
where \(\mathcal H\) is the chosen linear metric-perturbation distribution
space.  If \(\xi\) is the centered Gaussian generalized random tensor with
covariance \(N\), then \(h=\mathcal G_{\rm ret}\xi\) is a centered Gaussian
random metric perturbation.  For any real linear observables
\(\ell_1,\ell_2\in\mathcal H^*\),
\begin{equation}
  \mathbb E[\ell_1(h)\ell_2(h)]
  =
  N(\mathcal G_{\rm ret}^*\ell_1,\mathcal G_{\rm ret}^*\ell_2),
  \label{eq:semiclassical-pushforward-covariance-abstract}
\end{equation}
where \(\mathcal G_{\rm ret}^*:\mathcal H^*\to\mathcal E\) is the transpose.
In kernel notation this is exactly
\eqref{eq:semiclassical-metric-covariance}.  The resulting covariance is
positive semidefinite on metric observables:
\begin{equation}
  \mathbb E[\ell(h)^2]\ge0 .
  \label{eq:semiclassical-metric-covariance-positivity}
\end{equation}
For a centered Gaussian generalized random tensor, the characteristic
functional on \(\mathcal E\) is
\begin{equation}
  \mathbb E[e^{\ii\xi(f)}]
  =
  \exp\!\left[-\frac12N(f,f)\right].
  \label{eq:semiclassical-noise-characteristic-functional}
\end{equation}
For \(\ell\in\mathcal H^*\),
\begin{equation}
  \ell(h)=\ell(\mathcal G_{\rm ret}\xi)
  =
  \xi(\mathcal G_{\rm ret}^*\ell).
  \label{eq:semiclassical-observable-pullback}
\end{equation}
Thus the finite collection
\((\ell_1(h),\ldots,\ell_n(h))\) has characteristic function
\begin{equation}
  \exp\!\left[
    -\frac12
    \sum_{a,b}t_at_b\,
    N(\mathcal G_{\rm ret}^*\ell_a,\mathcal G_{\rm ret}^*\ell_b)
  \right],
  \label{eq:semiclassical-pushforward-characteristic-function}
\end{equation}
which is Gaussian with covariance
\eqref{eq:semiclassical-pushforward-covariance-abstract}.  Positivity
follows by setting \(\ell_1=\ell_2=\ell\) and using positivity of \(N\).
When \(\ell\) is evaluation against a compactly supported test tensor on the
metric side, \(\mathcal G_{\rm ret}^*\ell\) is represented by the retarded
Green kernel in \eqref{eq:semiclassical-metric-covariance}, producing the
displayed kernel formula.

Only gauge-invariant or gauge-fixed metric observables should be inserted in
the pushforward covariance formula above.  If
\(\ell\) is pure gauge and the retarded inverse has been constructed after
gauge fixing, \(\ell(h)\) is a coordinate artifact, not a physical metric
fluctuation.  The formula is therefore a covariance statement for the declared
linearized problem, not a replacement for the gravitational constraint
analysis.

\begin{controlledapproximation}[Large-\(N_{\rm sp}\) semiclassical scaling
window]
\label{ca:semiclassical-large-n-scaling-window}
The stochastic equation is often justified as the first correction to a
large-species semiclassical limit, but the scaling is a statement about
stress-tensor cumulants and the gravitational coupling, not a slogan.  Take
\(N_{\rm sp}\) independent identical matter sectors in the same admissible
Hadamard state class.  Stress-tensor products are understood after smearing
against the retained compactly supported tensor tests and after using the fixed
renormalization prescription; no pointwise product of distribution-valued
tensors is being assumed.  Write
\begin{equation}
  T_{\mu\nu}^{(N_{\rm sp})}
  =
  \sum_{i=1}^{N_{\rm sp}}T_{\mu\nu}^{(i)},
  \qquad
  \langle T_{\mu\nu}^{(N_{\rm sp})}\rangle
  =
  N_{\rm sp}\,t_{\mu\nu},
  \qquad
  N^{(N_{\rm sp})}
  =
  N_{\rm sp}\,N_1 ,
  \label{eq:semiclassical-large-n-stress-noise}
\end{equation}
where \(N_1\) is the connected stress-noise kernel of one sector in the fixed
renormalization prescription.  Multiply the gravitational equation by
\(\kappa_N=8\pi G_N\) and keep
\begin{equation}
  \bar\kappa
  :=
  N_{\rm sp}\kappa_N
  \quad\text{fixed as}\quad
  N_{\rm sp}\to\infty .
  \label{eq:semiclassical-large-n-kappa-scaling}
\end{equation}
Then the mean source in the rescaled metric equation is finite,
\(\kappa_N\langle T^{(N_{\rm sp})}\rangle=\bar\kappa\,t\), while the source
covariance seen by the metric is
\begin{equation}
  \kappa_N^2 N^{(N_{\rm sp})}
  =
  \frac{\bar\kappa^2}{N_{\rm sp}}\,N_1 .
  \label{eq:semiclassical-large-n-source-noise}
\end{equation}
Thus, on a retained response window with unit-normalized inverse bound
\(M_{\rm cl}\), the metric covariance satisfies
\begin{equation}
  \operatorname{tr}\mathsf C_h^{(N_{\rm sp})}
  \le
  \frac{\bar\kappa^2M_{\rm cl}^2}{N_{\rm sp}}\,
  \operatorname{tr}\mathsf N_1 ,
  \label{eq:semiclassical-large-n-metric-noise}
\end{equation}
where \(\mathsf N_1\) is the retained one-sector noise matrix.  For retained
smeared stress observables \(T_a=T(f_a)\), the same source-cumulant counting
gives, for every \(r\ge2\),
\begin{equation}
  \kappa_N^r
  \operatorname{Cum}_r
  \bigl(T_{a_1}^{(N_{\rm sp})},\ldots,T_{a_r}^{(N_{\rm sp})}\bigr)
  =
  \frac{\bar\kappa^r}{N_{\rm sp}^{\,r-1}}\,
  \operatorname{Cum}_r
  \bigl(T_{a_1}^{(1)},\ldots,T_{a_r}^{(1)}\bigr) .
  \label{eq:semiclassical-large-n-cumulants}
\end{equation}
Equation \eqref{eq:semiclassical-large-n-cumulants} is a statement about
rescaled stress-source cumulants.  The Gaussian Einstein--Langevin equation
uses only a centered Gaussian stochastic source with covariance
\(N^{(N_{\rm sp})}\), so it constructs the leading induced two-point metric
covariance, namely the linear retarded pushforward
\eqref{eq:semiclassical-large-n-metric-noise}, around the mean geometry.  Its
rms metric size is \(O(N_{\rm sp}^{-1/2})\).  Higher connected metric cumulants
require extra data: either a non-Gaussian stochastic hierarchy or the
higher-order closed-time-path influence functional, the retarded pushforward of
the smeared higher stress cumulants, and the nonlinear gravitational response
terms at the same order.  The \(N_{\rm sp}^{1-r}\) source scaling is input for
that separate construction, not a result of the Gaussian equation by itself.
These statements require the response bound, the one-sector renormalized stress
cumulants, and the state-transport residuals to stay uniform in \(N_{\rm sp}\).
They also require independence or sufficiently weak cross-correlations between
sectors.  If pairwise sector correlations contribute \(N_{12}\), the noise
contains \(N_{\rm sp}(N_{\rm sp}-1)N_{12}\); after multiplying by
\(\kappa_N^2\), this term is not \(1/N_{\rm sp}\)-suppressed.  Likewise,
keeping \(G_N\) fixed makes both the mean geometry and the metric noise grow
with \(N_{\rm sp}\).  The
large-species argument also does not remove the gravitational EFT cutoff or
species-cutoff constraint: curvature invariants and retained response
frequencies or momenta must remain small compared with the microscopic and
gravitational cutoff scales of the chosen hybrid theory.  Equivalently, the
curvature radius and the response wavelengths must remain large compared with
the corresponding cutoff lengths.
\end{controlledapproximation}

\section{Finite Response Window And Stability Diagnostics}

The preceding formulae become physically useful only after one asks whether
the retarded inverse is bounded in the sector being tested.  In a stationary
background, after gauge fixing and after solving or projecting out the
linearized constraints, let \(e_a\), \(a=1,\ldots,n\), be a finite set of
real metric perturbation modes or gauge-invariant metric observables.  Let
\(f_a\), \(a=1,\ldots,n\), be the corresponding finite set of real smeared
stress-tensor sources.  The retained linearized equation in frequency space
has the form
\begin{equation}
  \mathsf D_{ab}(\omega)h_b(\omega)=j_a(\omega),
  \qquad
  \mathsf D(\omega)
  =
  \mathsf K(\omega)-\frac12\mathsf\Pi^{\rm ret}(\omega),
  \label{eq:semiclassical-retained-response-matrix}
\end{equation}
where \(\mathsf K\) is the gravitational and local-contact part of the
linearized operator and \(\mathsf\Pi^{\rm ret}\) is the projected retarded
stress-tensor commutator.  The source \(j\) may include a state perturbation,
an external classical stress perturbation, or the stochastic source in the
Einstein--Langevin approximation.  All factors of \(G_N,\alpha,\beta\) are
included in \(\mathsf K\); equivalently one may multiply the whole equation
by \(8\pi G_N\), in which case the same test is applied to the rescaled
matrix.

\begin{controlledapproximation}[Finite semiclassical response window]
\label{ca:semiclassical-finite-response-window}
Fix a frequency interval \(I\), a retained metric-source sector as above, and
a reduced low-energy gravitational operator in which the cutoff-scale
higher-derivative roots have been removed by the EFT prescription.  Suppose
that the projected retarded response has no pole in the upper half of the
complex frequency plane in the strip obtained by continuing \(I\), and that
\begin{equation}
  M_I
  :=
  \sup_{\omega\in I}
  \left\lVert \mathsf D(\omega+\ii0)^{-1}\right\rVert_2
  <\infty .
  \label{eq:semiclassical-response-window-bound}
\end{equation}
Then, for every retained source vector \(j(\omega)\),
\begin{equation}
  \lVert h(\omega)\rVert_2
  \le
  M_I\,\lVert j(\omega)\rVert_2,
  \qquad
  \omega\in I.
  \label{eq:semiclassical-mean-response-bound}
\end{equation}
If the retained stress-tensor noise matrix
\[
  \mathsf N_{ab}(\omega)
  =
  N(f_a(\omega),f_b(\omega))
\]
is positive semidefinite, then the induced retained metric covariance is
\begin{equation}
  \mathsf C_h(\omega)
  =
  \mathsf D(\omega+\ii0)^{-1}
  \mathsf N(\omega)
  \mathsf D(\omega+\ii0)^{-*},
  \label{eq:semiclassical-retained-metric-covariance}
\end{equation}
and it obeys
\begin{equation}
  v^*\mathsf C_h(\omega)v\ge0,
  \qquad
  \operatorname{tr}\mathsf C_h(\omega)
  \le
  M_I^2\,\operatorname{tr}\mathsf N(\omega).
  \label{eq:semiclassical-retained-noise-bound}
\end{equation}
The retained semiclassical approximation is small in this window only if the
dimensionless quantities
\begin{equation}
  \varepsilon_{\rm mean}
  =
  \sup_{\omega\in I}
  \frac{\lVert h_{\rm mean}(\omega)\rVert_2}{h_{\rm ref}},
  \qquad
  \varepsilon_{\rm fluc}^2
  =
  \sup_{\omega\in I}
  \frac{\operatorname{tr}\mathsf C_h(\omega)}{h_{\rm ref}^2}
  \label{eq:semiclassical-response-small-parameters}
\end{equation}
are both much smaller than one, and if omitted modes and state changes remain
smaller than the same reference scale \(h_{\rm ref}\).  Here \(h_{\rm ref}\)
is fixed by the curvature radius or by the amplitude at which the chosen
linearized metric chart ceases to be valid.
\end{controlledapproximation}

The proof is finite-dimensional linear algebra, but the diagnostic is not a
mere matrix exercise.  Upper-half-plane poles of \(\mathsf D^{-1}\) are
linear instabilities or runaway roots of the retained semiclassical problem;
large \(M_I\) means that a small stress-tensor perturbation is amplified into
a large metric response.  A positive noise kernel then gives a positive
metric covariance by pushforward, and
\[
  \operatorname{tr}(\mathsf D^{-1}\mathsf N\mathsf D^{-*})
  =
  \operatorname{tr}(\mathsf D^{-*}\mathsf D^{-1}\mathsf N)
  \le
  \lVert \mathsf D^{-1}\rVert_2^2\operatorname{tr}\mathsf N ,
\]
because \(0\le \mathsf D^{-*}\mathsf D^{-1}\le
\lVert\mathsf D^{-1}\rVert_2^2\mathbf 1\) on the retained sector.  This is
the bridge from the formal semiclassical equation to a physical
backreaction statement: the mean response, the fluctuation response, and the
absence of unstable response poles must be checked together.

\section{Worked Retained Potential-Source Coordinate}
\label{sec:semiclassical-worked-interacting-source-shift}

The response-window criterion is empty unless the source vector is supplied
by a declared stress-tensor construction.  The following calculation is
deliberately narrower than a full interacting stress tensor.  It isolates the
explicit potential insertion from the scalar interaction of
Chapter~\ref{ch:curved-perturbative-algebraic-qft}, evaluates that insertion
in the reference free quasifree state, and then records how this retained
source coordinate would enter the finite response matrix.  Work to first
order in \(\lambda\), fix a Hadamard Wick coordinate \(H\), and assume a
centered quasifree Hadamard state \(\omega\) on a stationary region \(U\)
where the smooth Wick variance
\begin{equation}
  \Sigma_{\omega,H}(x)
  :=
  \omega\!\left(\Phi_H^2(x)\right)
  \label{eq:semiclassical-wick-variance}
\end{equation}
is constant.  The interaction cutoff is taken to be \(1\) on \(U\); edges of
the cutoff are outside the observables considered here.  The potential
insertion in the first-order variation of the interaction action gives
\begin{equation}
  \delta_\lambda T_{\mu\nu}^{\rm pot}(x)
  =
  -g_{\mu\nu}(x)\frac{\lambda}{4!}\Phi_H^4(x).
  \label{eq:semiclassical-interacting-potential-stress}
\end{equation}
For a centered quasifree state,
\[
  \omega\!\left(\Phi_H^4(x)\right)
  =
  3\Sigma_{\omega,H}(x)^2,
\]
so the observable output for any test tensor \(f^{\mu\nu}\) supported in
\(U\) is
\begin{equation}
  \delta_\lambda
  \omega\!\left(T^{\rm pot}(f)\right)
  =
  -\frac{\lambda}{8}
  \int_U\dd\operatorname{vol}_g\,
  f^{\mu\nu}(x)g_{\mu\nu}(x)\Sigma_{\omega,H}^2 .
  \label{eq:semiclassical-interacting-smeared-source}
\end{equation}
Equivalently, the local retained potential source coordinate is
\begin{equation}
  \delta_\lambda
  \langle T_{\mu\nu}\rangle_{\omega,H}^{\rm pot}
  =
  -\rho_\lambda g_{\mu\nu},
  \qquad
  \rho_\lambda
  =
  \frac{\lambda}{8}\Sigma_{\omega,H}^2 .
  \label{eq:semiclassical-interacting-vacuum-source}
\end{equation}
Equation~\eqref{eq:semiclassical-interacting-vacuum-source} is not, by
itself, the full order-\(\lambda\) expectation value of the renormalized
interacting stress tensor.  A full first-order source has the decomposition
\begin{equation}
\begin{aligned}
  \delta_\lambda
  \langle T_{\mu\nu}^{\rm ren}\rangle_{\omega,H}^{\rm full}
  =
  &\ \delta_\lambda
  \langle T_{\mu\nu}\rangle_{\omega,H}^{\rm pot}
  +\delta_\lambda
  \left\langle R_V(T_{\mu\nu}^{\rm free})\right\rangle_{\omega,H}^{(1)}
  \\
  &\ +C_{\mu\nu}^{\rm int}[g;m,\xi,\lambda,H] .
\end{aligned}
  \label{eq:semiclassical-full-interacting-source-ledger}
\end{equation}
The second term is the Bogoliubov/interacting-field correction to the free
stress tensor, and \(C_{\mu\nu}^{\rm int}\) collects the chosen
time-ordered-product contact terms, interaction-dependent composite-field
counterterms, and local stress-tensor curvature terms.  The conservation
condition
\[
  \nabla^\mu
  \delta_\lambda
  \langle T_{\mu\nu}^{\rm ren}\rangle_{\omega,H}^{\rm full}
  =
  0
\]
is a Ward condition on this whole sum, not on the potential insertion alone.
This is the curved-spacetime pAQFT renormalization problem isolated by
Hollands--Wald local Wick products, time-ordered products, and interacting
stress-tensor conservation
\cite{HollandsWald2001LocalWick,HollandsWald2005Conservation,HollandsWald2015QFCS}.

The displayed \(\rho_\lambda\) is therefore a renormalized composite-field
coordinate inside a restricted sector.  Under a pure Hadamard coordinate
transport \(H\to H'=H+w\), the observable and state must be transported
together and the expectation value is unchanged.  A c-number change of the
displayed density instead compares two local covariant Wick prescriptions in
the same physical state.  If that finite Wick-square renormalization adds the
local scalar \(c\), then \(\Sigma_{\omega,H}\) is replaced in this comparison
by \(\Sigma_{\omega,H}+\hbar c\), and
\begin{equation}
  \rho_\lambda^{(c)}
  =
  \frac{\lambda}{8}
  \left(\Sigma_{\omega,H}+\hbar c\right)^2 .
  \label{eq:semiclassical-interacting-source-scheme-shift}
\end{equation}
This Wick-square-generated shift is only one finite-renormalization
subfamily.  A general locally covariant \(\Phi^4\) prescription also permits
independent local lower-field and geometric terms of the form
\[
  \Phi_H^4
  \mapsto
  \Phi_H^4
  +A_1m^2\Phi_H^2
  +A_2R\Phi_H^2
  +A_3m^4
  +A_4m^2R
  +A_5R^2
  +A_6R_{\alpha\beta}R^{\alpha\beta}
  +A_7\Box R ,
\]
subject to the chosen local covariance, scaling, and action-Ward conditions.
The stress tensor has its own independent conserved local curvature tensors,
as in Proposition~\ref{prop:finite-local-freedom-scalar-stress-tensor}, plus
the interaction-dependent contact terms in
\eqref{eq:semiclassical-full-interacting-source-ledger}.  Fixing the
Wick-square prescription by itself therefore does not make the number
\(\rho_\lambda\) a physical interacting energy-density prediction.

In the homogeneous retained-coordinate calculation,
\eqref{eq:semiclassical-interacting-vacuum-source} is conserved and may be
absorbed into a potential-sector gravitational coordinate
\begin{equation}
  \Lambda_{\rm eff}^{\rm pot}
  =
  \Lambda
  +
  8\pi G_N\rho_\lambda
  =
  \Lambda+\pi G_N\lambda\Sigma_{\omega,H}^2 .
  \label{eq:semiclassical-interacting-lambda-shift}
\end{equation}
The sign follows from the convention
\(E_{\mu\nu}^{\rm grav}=\langle T_{\mu\nu}\rangle\): a positive potential
energy gives \(T_{\mu\nu}=-\rho_\lambda g_{\mu\nu}\) and therefore moves the
source to the left side as \(+\rho_\lambda g_{\mu\nu}\).  For a physical
semiclassical equation this scalar coordinate must be replaced by the
corresponding scalar component of the conserved, scheme-fixed full source in
\eqref{eq:semiclassical-full-interacting-source-ledger}.

The same retained potential coordinate can be fed into the response
diagnostic as a toy source-coordinate check.  With the finite source basis
\(f_a\) of
\eqref{eq:semiclassical-retained-response-matrix}, define
\begin{equation}
  q_a
  =
  \int_U\dd\operatorname{vol}_g\, f_a^{\mu\nu}g_{\mu\nu},
  \qquad
  j_a^{(\lambda,{\rm pot})}
  =
  -\rho_\lambda q_a .
  \label{eq:semiclassical-interacting-retained-source}
\end{equation}
If the response-window hypotheses hold, then
\begin{equation}
  \lVert h_\lambda^{\rm pot}(\omega)\rVert_2
  \le
  M_I|\rho_\lambda|\lVert q\rVert_2,
  \qquad
  \omega\in I .
  \label{eq:semiclassical-interacting-response-bound}
\end{equation}
This is the retained-sector bookkeeping test: an interaction, a state
assumption, a renormalization coordinate, an observable potential-insertion
output, and a metric-response bound.  It becomes a physical backreaction
datum only after
\(\delta_\lambda\langle T_{\mu\nu}\rangle^{\rm pot}\) is replaced by the
conserved full source
\(\delta_\lambda
\langle T_{\mu\nu}^{\rm ren}\rangle_{\omega,H}^{\rm full}\).  If
\(\Sigma_{\omega,H}\) varies across the region, the potential term alone has
\(\nabla^\mu(-\rho_\lambda g_{\mu\nu})=-\partial_\nu\rho_\lambda\).  The
full interacting stress tensor must then include the derivative and local
counterterm pieces fixed by the same pAQFT renormalization prescription; the
potential-only display is no longer a conserved standalone source.

\begin{controlledapproximation}[Retained interacting potential noise]
\label{ca:semiclassical-interacting-potential-noise}
Keep the same Hadamard Wick coordinate, quasifree state, and potential
insertion as in
Section~\ref{sec:semiclassical-worked-interacting-source-shift}, but now ask
for the stress-tensor fluctuation sourced by the interaction.  Let
\[
  R_{\omega,H}(x,y)
  =
  \omega_2(x,y)-H_+(x,y),
  \qquad
  W_\omega(x,y)=\omega_2(x,y)
\]
Here \(H_+\) is the Hadamard two-point parametrix for the Wick chart, and
\[
  \Sigma_{\omega,H}(x)=R_{\omega,H}(x,x)
\]
is the smooth Wick-square expectation.  The separated cross contraction is the
full \(W_\omega(x,y)\); the smooth remainder \(R_{\omega,H}(x,y)\) is only one
piece of the typed \(H\)-chart calculation:
\begin{equation}
\begin{aligned}
  &\omega\!\left(
    \exp_H(t\Phi(x))\star_H\exp_H(s\Phi(y))
  \right) \\
  &\qquad =
  \exp\!\left(t s H_+(x,y)\right)
  \exp\!\left(
    \frac{t^2}{2}\Sigma_{\omega,H}(x)
    +t s R_{\omega,H}(x,y)
    +\frac{s^2}{2}\Sigma_{\omega,H}(y)
  \right) \\
  &\qquad =
  \exp\!\left(
    \frac{t^2}{2}\Sigma_{\omega,H}(x)
    +t s W_\omega(x,y)
    +\frac{s^2}{2}\Sigma_{\omega,H}(y)
  \right).
\end{aligned}
  \label{eq:semiclassical-star-H-state-wick-generating}
\end{equation}
The first exponential is the \(\star_H\)-product contraction, while the second
is the state evaluation of the \(H\)-Wick exponential.  They combine to the full
state two-point function between distinct insertions.  Differentiating
\eqref{eq:semiclassical-star-H-state-wick-generating} gives, away from the
diagonal and before the contact extension of the stress-tensor square,
\begin{equation}
\begin{aligned}
  &\omega\!\left(
    \Phi_H^4(x)\star_H\Phi_H^4(y)
  \right)_{\rm conn}
  \\
  &\qquad =
  72\,
  \Sigma_{\omega,H}(x)\Sigma_{\omega,H}(y)
  W_\omega(x,y)^2
  +
  24\,W_\omega(x,y)^4 .
\end{aligned}
  \label{eq:semiclassical-wick-four-noise-contractions}
\end{equation}
The disconnected term
\(9\Sigma_{\omega,H}(x)^2\Sigma_{\omega,H}(y)^2\) has been subtracted.  The
first term in
\eqref{eq:semiclassical-wick-four-noise-contractions} is present whenever
the state covariance differs from the Wick coordinate at the individual
insertions.  If the Wick coordinate is chosen from the same quasifree state,
then \(\Sigma_{\omega,H}=0\) and \(R_{\omega,H}=0\), but the separated
connected fourth-Wick-power covariance remains
\(24W_\omega(x,y)^4\).  A formula built only from the smooth remainder would
erase this propagation term.

Thus the retained potential contribution to the stress-noise kernel is
\begin{equation}
  N_{\mu\nu,\rho\sigma}^{(\lambda,{\rm pot})}(x,y)
  =
  \lambda^2 g_{\mu\nu}(x)g_{\rho\sigma}(y)
  \operatorname{Re}\!\left[
    \frac18\Sigma_{\omega,H}(x)\Sigma_{\omega,H}(y)
    W_\omega(x,y)^2
    +
    \frac1{24}W_\omega(x,y)^4
  \right],
  \label{eq:semiclassical-potential-noise-kernel}
\end{equation}
in the retained potential sector.  The real part is taken after the square and
fourth power; it is the symmetrized Lorentzian noise operation, not a
replacement of \(W_\omega(x,y)\) by \(\operatorname{Re}W_\omega(x,y)\) before
the polynomial is formed.  For the finite source basis
\eqref{eq:semiclassical-retained-response-matrix}, set
\[
  \mathsf N_{ab}^{(\lambda,{\rm pot})}
  =
  \int_{U\times U}
  q_a(x)
  \operatorname{Re}\!\left[
    \frac18\Sigma_{\omega,H}(x)\Sigma_{\omega,H}(y)
    W_\omega(x,y)^2
    +
    \frac1{24}W_\omega(x,y)^4
  \right]
  q_b(y)\,
  \lambda^2\,
  \dd\operatorname{vol}_g(x)\dd\operatorname{vol}_g(y),
\]
where \(q_a(x)=f_a^{\mu\nu}(x)g_{\mu\nu}(x)\).  This matrix is positive
semidefinite because it is the covariance of the retained random variables
\(-(\lambda/4!)\int q_a\Phi_H^4\).  The potential-noise contribution to the
retained metric covariance is therefore
\begin{equation}
  \mathsf C_h^{(\lambda,{\rm pot})}
  =
  \mathsf D^{-1}\mathsf N^{(\lambda,{\rm pot})}\mathsf D^{-*},
  \qquad
  \operatorname{tr}\mathsf C_h^{(\lambda,{\rm pot})}
  \le
  M_I^2\operatorname{tr}\mathsf N^{(\lambda,{\rm pot})}.
  \label{eq:semiclassical-potential-noise-response-bound}
\end{equation}

The full interacting noise kernel also contains cross terms with the
Bogoliubov correction to the free stress tensor, local contact extensions of
stress-tensor products, derivative terms, and finite composite-operator
counterterms.  Equation~\eqref{eq:semiclassical-potential-noise-kernel} is
therefore the fluctuation analogue of
\eqref{eq:semiclassical-interacting-vacuum-source}: it is a retained
potential-sector coordinate.  It becomes a physical Einstein--Langevin input
only after the conserved full stress tensor and its renormalized square have
been fixed in the same locally covariant scheme.
\end{controlledapproximation}

\section{Ward Diagnostics For Interacting Sources}
\label{sec:semiclassical-interacting-source-ward-diagnostics}

The retained potential source and retained potential noise above are useful only
after their conservation status has been made explicit.
Section~\ref{sec:paqft-local-coupling-ward-ledger} gives the upstream pAQFT
version of the same issue: a nonconstant switching function is an external
coupling source, not a conserved stress tensor by itself.  In a region where
the Wick-square coordinate varies,
\begin{equation}
  s_\nu^{\rm pot}
  :=
  \nabla^\mu
  \delta_\lambda\langle T_{\mu\nu}\rangle_{\omega,H}^{\rm pot}
  =
  -\partial_\nu\rho_\lambda .
  \label{eq:semiclassical-potential-source-divergence}
\end{equation}
The missing term is therefore part of the source equation.  The interacting
correction in
\eqref{eq:semiclassical-full-interacting-source-ledger} must satisfy
\begin{equation}
  \nabla^\mu
  \left(
    \delta_\lambda
    \left\langle R_V(T_{\mu\nu}^{\rm free})\right\rangle_{\omega,H}^{(1)}
    +
    C_{\mu\nu}^{\rm int}[g;m,\xi,\lambda,H]
  \right)
  =
  \partial_\nu\rho_\lambda ,
  \label{eq:semiclassical-interacting-ward-diagnostic-source-equation}
\end{equation}
up to the same distributional contact terms and field-equation ideals used in
the locally covariant construction of the stress tensor.  This is where the
physics of the interacting stress tensor enters the backreaction problem: the
potential energy coordinate supplies one part of the source, but the
Bogoliubov correction, time-ordered-product extension, and local finite
counterterms determine whether the source can actually drive a metric.

The finite response window of
Controlled Approximation~\ref{ca:semiclassical-finite-response-window} gives a
compact way to keep this requirement visible.  Choose retained stress profiles
\(\tau_{\mu\nu}^{(a)}\) dual to the source coordinates \(j_a\), and test the
divergence against compact vector fields \(u_\alpha^\nu\).  This gives the
stress-divergence Ward map
\begin{equation}
  (\mathsf B_T j)_\alpha
  =
  \sum_a j_a
  \int_U
  u_\alpha^\nu\nabla^\mu\tau_{\mu\nu}^{(a)}
  \,\dd\operatorname{vol}_g .
  \label{eq:semiclassical-retained-ward-map}
\end{equation}
The subscript is part of the retained datum:
\(\mathsf B_T:V_T\to V_{\nabla T}\) maps stress-source coordinates to
divergence tests.  It is not the metric gauge generator, not the gravitational
constraint map, and not a right-null map for the retarded stress response.
A retained source is admissible in the conserved backreaction sector only if
\begin{equation}
  \mathsf B_T j^{\rm full}=0,
  \qquad
  j^{\rm full}
  =
  j^{(\lambda,{\rm pot})}
  +
  j^{(\lambda,{\rm Bog})}
  +
  j^{(\lambda,{\rm ct})}.
  \label{eq:semiclassical-retained-source-ward-condition}
\end{equation}
The retained Ward equation fixes the longitudinal part of the missing source
coordinate.  In orthonormal retained coordinates, when \(\mathsf B_T\) has full
row rank on the tested constraint subspace, one convenient particular
longitudinal repair of a raw source coordinate is
\begin{equation}
  c_{\rm min}
  =
  -
  \mathsf B_T^*
  \left(\mathsf B_T\mathsf B_T^*\right)^{-1}
  \mathsf B_T j^{\rm raw},
  \qquad
  \mathsf P_{\ker \mathsf B_T}
  =
  \mathsf I-\mathsf B_T^*
  \left(\mathsf B_T\mathsf B_T^*\right)^{-1}
  \mathsf B_T,
  \qquad
  j_{\rm min}
  =
  \mathsf P_{\ker \mathsf B_T}j^{\rm raw}.
  \label{eq:semiclassical-retained-ward-projector}
\end{equation}
With non-orthonormal retained test functions the same formula is written with
the corresponding Gram matrices inserted.  Equation
\eqref{eq:semiclassical-retained-ward-projector} is a retained-sector
diagnostic.  It detects a longitudinal violation such as
\(\mathsf B_T j^{(\lambda,{\rm pot})}\ne0\), records a least-norm particular
repair of that violation, and exposes the transverse datum that the Ward
identity leaves open:
\begin{equation}
  j^{\rm full}
  =
  j_{\rm min}+k,
  \qquad
  k\in\ker\mathsf B_T .
  \label{eq:semiclassical-retained-ward-transverse-freedom}
\end{equation}
The vector \(k\) is part of the interacting QFT data.  It is fixed, in the
full construction, by the Bogoliubov correction, time-ordered-product
extensions, composite-operator finite renormalizations, local covariance, action
Ward identities, and physical renormalization conditions.  The Euclidean
least-norm retained vector \(j_{\rm min}\) is therefore a diagnostic or an
explicit model choice inside the retained source space; the physical source is
the \(j^{\rm full}\) obtained from the pAQFT construction and then tested by
\(\mathsf B_T j^{\rm full}=0\).
In the homogeneous special case \(\partial_\nu\rho_\lambda=0\), the retained
potential coordinate is already in \(\ker\mathsf B_T\), which is why the
cosmological-coordinate shift in
\eqref{eq:semiclassical-interacting-lambda-shift} is a consistent scalar
bookkeeping coordinate there.

The same conservation test applies to fluctuations.  The full interacting
noise kernel must obey the distributional Ward identities
\[
  \nabla_x^\mu N_{\mu\nu,\rho\sigma}^{\rm full}(x,y)=0,
  \qquad
  \nabla_y^\rho N_{\mu\nu,\rho\sigma}^{\rm full}(x,y)=0,
\]
after the renormalized square of the interacting stress tensor has been fixed.
In the retained sector this says
\begin{equation}
  \mathsf B_T\mathsf N^{\rm full}=0,
  \qquad
  \mathsf N^{\rm full}\mathsf B_T^*=0 .
  \label{eq:semiclassical-retained-noise-ward-condition}
\end{equation}
The full covariance includes all component covariances and cross-covariances:
\[
  \mathsf N^{\rm full}
  =
  \mathsf N^{\rm pot}
  +
  \mathsf N^{\rm Bog}
  +
  \mathsf N^{\rm ct}
  +
  \mathsf N^{\rm cross}.
\]
If this full object has already been constructed and satisfies
\eqref{eq:semiclassical-retained-noise-ward-condition}, then
\begin{equation}
  \mathsf P_{\ker \mathsf B_T}
  \mathsf N^{\rm full}
  \mathsf P_{\ker \mathsf B_T}^*
  =
  \mathsf N^{\rm full}.
  \label{eq:semiclassical-retained-noise-ward-projection}
\end{equation}
Applying the projection to a partial covariance, such as
\(\mathsf N^{(\lambda,{\rm pot})}\), gives a positive diagnostic covariance for
the chosen least-norm retained model.  It also omits the interacting correction
fluctuations and the raw/correction cross-correlations that enter
\(\mathsf N^{\rm full}\).  The physical retained metric data are consequently
\begin{equation}
  h^{\rm full}(\omega)
  =
  \mathsf D(\omega+\ii0)^{-1}j^{\rm full}(\omega),
  \qquad
  \mathsf C_h^{\rm full}(\omega)
  =
  \mathsf D(\omega+\ii0)^{-1}
  \mathsf N^{\rm full}(\omega)
  \mathsf D(\omega+\ii0)^{-*},
  \label{eq:semiclassical-ward-clean-response-data}
\end{equation}
with the same bounded-response estimates as in
\eqref{eq:semiclassical-mean-response-bound} and
\eqref{eq:semiclassical-retained-noise-bound}, provided the Ward map, the
constraint-reduced source space, and the gauge-fixed gravitational inverse
\(\mathsf D^{-1}\) are defined on the same retained sector.  This is the point
of keeping the potential-sector calculation narrow: it is a computable
coordinate inside the interacting source, while the conserved mean and noise
that feed semiclassical gravity require the full Ward-compatible pAQFT package.

\begin{controlledapproximation}[Homogeneous FLRW interacting source and noise
closure]
\label{ca:semiclassical-flrw-interacting-source-noise-closure}
The conservation issue becomes physical in the simplest cosmological retained
sector.  On a spatially flat FLRW metric
\[
  ds^2=-dt^2+a(t)^2d\mathbf x^2,
  \qquad H=\frac{\dot a}{a},
\]
write a homogeneous stress perturbation as
\begin{equation}
  \delta T^\mu{}_\nu
  =
  \operatorname{diag}(-\delta\rho,\delta p,\delta p,\delta p).
  \label{eq:semiclassical-flrw-homogeneous-source}
\end{equation}
The linearized Bianchi identity is
\begin{equation}
  \delta\dot\rho
  +
  3H(\delta\rho+\delta p)
  +
  3\delta H(\rho_0+p_0)
  =
  0 ,
  \label{eq:semiclassical-flrw-linear-ward}
\end{equation}
where \((\rho_0,p_0)\) is the background matter source.  In a de Sitter or
vacuum-energy retained cell, \(\rho_0+p_0=0\), and the last term drops.  The
potential coordinate
\eqref{eq:semiclassical-interacting-vacuum-source} has
\[
  \delta\rho_{\rm pot}=\rho_\lambda(t),
  \qquad
  \delta p_{\rm pot}=-\rho_\lambda(t),
  \qquad
  \rho_\lambda(t)=\frac{\lambda}{8}\Sigma_{\omega,H}(t)^2 .
\]
It is a conserved cosmological-constant coordinate only when the transported
state and Wick chart make \(\dot\rho_\lambda=0\) on the retained window.  If
\(\Sigma_{\omega,H}(t)\) changes because the state is being transported
through an expanding background, the potential coordinate by itself has
conservation residual \(\dot\rho_\lambda\).  The missing interacting pieces
must supply a correction
\((\delta\rho_{\rm corr},\delta p_{\rm corr})\) satisfying
\begin{equation}
  \dot\rho_\lambda
  +
  \delta\dot\rho_{\rm corr}
  +
  3H(\delta\rho_{\rm corr}+\delta p_{\rm corr})
  +
  3\delta H(\rho_0+p_0)
  =
  0 .
  \label{eq:semiclassical-flrw-interacting-ward-closure}
\end{equation}
When \(H\ne0\), this may be written as the pressure closure
\begin{equation}
  \delta p_{\rm corr}
  =
  -\delta\rho_{\rm corr}
  -
  \frac{
    \dot\rho_\lambda+\delta\dot\rho_{\rm corr}
    +3\delta H(\rho_0+p_0)
  }{3H}.
  \label{eq:semiclassical-flrw-pressure-closure}
\end{equation}
At a bounce or at a turning point with \(H=0\), one should use the unsolved
equation \eqref{eq:semiclassical-flrw-interacting-ward-closure}; the pressure
formula is only a coordinate expression in an expanding or contracting chart.

For a flat de Sitter retained cell, the two linear metric equations are
\begin{equation}
  6H\,\delta H
  =
  \kappa\,\delta\rho,
  \qquad
  -2\,\delta\dot H
  =
  \kappa(\delta\rho+\delta p),
  \qquad
  \kappa=8\pi G_N .
  \label{eq:semiclassical-flrw-linear-friedmann-raychaudhuri}
\end{equation}
They are compatible only for sources obeying
\eqref{eq:semiclassical-flrw-linear-ward}.  A time-dependent
\(\rho_\lambda(t)\) inserted with \(p=-\rho_\lambda\) gives a changing
Friedmann source but a vanishing Raychaudhuri source, so the two metric
equations define incompatible responses.  After the Ward closure
\eqref{eq:semiclassical-flrw-interacting-ward-closure}, differentiating the
Friedmann equation gives the Raychaudhuri equation.  The pressure and
state-transport terms are therefore not optional refinements of the potential
energy calculation.  They are the part of the interacting stress tensor that
lets the metric respond consistently.

The same lesson applies to stochastic backreaction.  On a retained homogeneous
time cell, collect the stress-noise coordinates as
\[
  \zeta
  =
  (\xi_{\dot\rho},\xi_\rho,\xi_p)^T,
  \qquad
  \mathsf B_{\rm FLRW}
  =
  \begin{pmatrix}
    1&3H&3H
  \end{pmatrix}.
\]
Here \(\xi_{\dot\rho}\) is not an independent random coordinate.  It is the
distributional time derivative of the density noise in the same transported
state and renormalization scheme.  At the covariance level the retained kernel
must therefore satisfy, before the algebraic Ward test is applied,
\begin{equation}
  N_{\dot\rho,\rho}(t,t')
  =
  \partial_t N_{\rho,\rho}(t,t'),
  \qquad
  N_{\rho,\dot\rho}(t,t')
  =
  \partial_{t'} N_{\rho,\rho}(t,t'),
  \qquad
  N_{\dot\rho,\dot\rho}(t,t')
  =
  \partial_t\partial_{t'}N_{\rho,\rho}(t,t').
  \label{eq:semiclassical-flrw-derivative-noise-consistency}
\end{equation}
Equivalently, in a finite time cell the rows and columns involving
\(\xi_{\dot\rho}\) are obtained by pushing the density covariance through the
retained derivative matrix.  A positive covariance for three formal variables
\((\xi_{\dot\rho},\xi_\rho,\xi_p)\) is not yet a physical stress-noise
covariance if these derivative identities fail.
The Einstein--Langevin source covariance in this sector must satisfy
\begin{equation}
  \mathsf B_{\rm FLRW}\mathsf N_\zeta=0,
  \qquad
  \mathsf N_\zeta\mathsf B_{\rm FLRW}^*=0 .
  \label{eq:semiclassical-flrw-noise-ward-condition}
\end{equation}
A covariance for the potential density alone, even if positive, does not
determine the pressure and time-derivative noise required by
\eqref{eq:semiclassical-flrw-noise-ward-condition}.  For Ward-clean noise, the
Friedmann and Raychaudhuri routes to the Hubble fluctuation agree:
\begin{equation}
  \xi_{\dot H}^{\rm F}
  =
  \frac{\kappa}{6H}\xi_{\dot\rho},
  \qquad
  \xi_{\dot H}^{\rm R}
  =
  -\frac{\kappa}{2}(\xi_\rho+\xi_p),
  \qquad
  \xi_{\dot H}^{\rm F}-\xi_{\dot H}^{\rm R}=0
  \quad\text{on }\ker\mathsf B_{\rm FLRW}.
  \label{eq:semiclassical-flrw-hubble-noise-closure}
\end{equation}
Thus a cosmological interacting backreaction calculation must provide the
transported state, the full renormalized stress tensor, its pressure component,
and the Ward-clean stress-noise covariance before the retained Hubble response
or Hubble noise is read as physics.  This is the interacting analogue of the
produced-stress continuity test in
Chapter~\ref{ch:curved-cosmological-spacetimes-particle-creation}.
\end{controlledapproximation}

\begin{example}[Finite retained Ward-completion laboratory]
\label{ex:semiclassical-retained-ward-completion-laboratory}
The warning above is not a formal nicety; it changes the metric source.  In a
two-coordinate retained sector, take the single Ward test
\[
  \mathsf B_T=(1,-1),
  \qquad
  \mathsf P_{\ker\mathsf B_T}
  =
  \frac12
  \begin{pmatrix}
    1&1\\
    1&1
  \end{pmatrix}.
\]
The coordinate \(j_1-j_2\) represents the tested longitudinal divergence of a
candidate stress source, for instance the retained image of
\(-\partial_\nu\rho_\lambda\) in
\eqref{eq:semiclassical-potential-source-divergence}.  If the potential-only
calculation gives
\[
  j^{\rm raw}
  =
  \rho_*
  \binom{3}{1},
  \qquad
  \mathsf B_T j^{\rm raw}=2\rho_*,
\]
then the least-norm Ward repair and the projected retained source are
\begin{equation}
  c_{\rm min}
  =
  \rho_*
  \binom{-1}{1},
  \qquad
  j_{\rm min}
  =
  \mathsf P_{\ker\mathsf B_T}j^{\rm raw}
  =
  \rho_*
  \binom{2}{2}.
  \label{eq:semiclassical-finite-ward-laboratory-source}
\end{equation}
This vector is conserved in the retained test, but it is not the physical
interacting stress source by itself.  The most general Ward-clean retained
source has the form
\[
  j^{\rm full}
  =
  j_{\rm min}
  +
  k\rho_*
  \binom{1}{1},
\]
where the coefficient \(k\) is fixed by the interacting stress-tensor
construction and its renormalization conditions.  A transverse finite
counterterm changes \(k\); it cannot cancel the longitudinal violation
\(\mathsf B_T j^{\rm raw}=2\rho_*\).  With a Ward-reduced inverse acting on the
conserved line, for example
\[
  \mathsf D^{-1}_{\rm red}
  =
  M_*
  \mathsf P_{\ker\mathsf B_T},
\]
the diagnostic retained response is
\[
  h_{\rm min}
  =
  2M_*\rho_*
  \binom{1}{1}.
\]
Feeding \(j^{\rm raw}\) to an unconstrained inverse would instead ask the
gravitational equation to respond to a nonconserved source.

The same laboratory exposes the fluctuation issue.  A partial potential-noise
matrix such as
\[
  \mathsf N^{\rm raw}
  =
  \rho_*^2
  \begin{pmatrix}
    5&1\\
    1&2
  \end{pmatrix}
\]
does not obey the retained Ward identities.  The projected diagnostic
covariance is
\begin{equation}
  \mathsf N_{\rm diag}
  =
  \mathsf P_{\ker\mathsf B_T}\mathsf N^{\rm raw}
  \mathsf P_{\ker\mathsf B_T}^*
  =
  \frac94\rho_*^2
  \begin{pmatrix}
    1&1\\
    1&1
  \end{pmatrix}.
  \label{eq:semiclassical-finite-ward-laboratory-noise}
\end{equation}
It is positive and Ward-clean, so it is a valid covariance for the explicit
least-norm retained model.  It is still not the full noise of the interacting
stress tensor.  Missing Bogoliubov, derivative, contact, and finite-mixing
terms can contribute an additional positive Ward-clean covariance
\(\mathsf N_{\rm miss}\) with
\(\mathsf B_T\mathsf N_{\rm miss}=0=\mathsf N_{\rm miss}\mathsf B_T^*\).  The
metric fluctuation budget must therefore use
\[
  \mathsf C_h^{\rm full}
  =
  \mathsf D^{-1}_{\rm red}
  \left(\mathsf N_{\rm diag}+\mathsf N_{\rm miss}\right)
  \mathsf D_{\rm red}^{-*},
\]
not only the projected partial covariance.  This is the finite retained
version of the physical rule used below: only the scheme-fixed interacting
stress tensor, its connected square, and its retarded response constitute a
backreaction input.
\end{example}

\begin{controlledapproximation}[Interacting stress-tensor and noise package]
\label{ca:semiclassical-interacting-stress-noise-package}
The retained Ward projector diagnoses admissible conserved coordinates; the
physical backreaction input comes from the interacting stress tensor itself.
Fix a pAQFT renormalization prescription \(\mathcal S\), an interaction \(V\),
a Hadamard state \(\omega\) in the selected state class, and a retained
stress-tensor test basis \(\tau_a^{\mu\nu}\).  A backreaction-ready interacting
package on this retained sector consists of the three objects
\begin{equation}
\begin{aligned}
  j_a^{\rm full}
  &=
  \omega\!\left(T_{\rm int}^{\mathcal S,V}(\tau_a)\right),\\
  \mathsf N_{ab}^{\rm full}
  &=
  \frac12\,
  \omega\!\left(
    \Delta T_{\rm int}^{\mathcal S,V}(\tau_a)
    \star_{\mathcal S,V}
    \Delta T_{\rm int}^{\mathcal S,V}(\tau_b)
    +
    \Delta T_{\rm int}^{\mathcal S,V}(\tau_b)
    \star_{\mathcal S,V}
    \Delta T_{\rm int}^{\mathcal S,V}(\tau_a)
  \right),\\
  \mathsf K_{ab}^{\rm ret}
  &=
  \ii\,\omega\!\left(
    \left[
      T_{\rm int}^{\mathcal S,V}(\tau_a),
      T_{\rm int}^{\mathcal S,V}(\tau_b)
    \right]
  \right)_{\rm ret}
  +\mathsf K_{ab}^{\rm loc}.
\end{aligned}
  \label{eq:semiclassical-full-interacting-package}
\end{equation}
Here \(\Delta T=T-\omega(T)\), and
\(\star_{\mathcal S,V}\) is the unordered interacting algebra product in the
same scheme in which \(T_{\rm int}^{\mathcal S,V}\) is constructed.  The
retarded response \(\mathsf K^{\rm ret}\) may contain local
metric-variation/contact terms because it is a response/effective-action
object.  The noise covariance \(\mathsf N^{\rm full}\) is not given an
independent time-ordered contact parameter.  Any local term in the interacting
noise must be derived from the chosen stress observable and unordered product,
and must then pass positivity and Ward tests as a covariance.  The same
renormalized operator-valued distribution \(T_{\rm int}^{\mathcal S,V}\) must
be used in all three entries.  Otherwise the mean equation, the
Einstein--Langevin noise, and the retarded response are coordinates from
different theories.

If, for bookkeeping, the interacting tensor is decomposed in one local chart as
\[
  T_{\rm int}^{\mathcal S,V}
  =
  T^{(0)}
  +T^{\rm pot}
  +T^{\rm Bog}
  +T^{\rm der}
  +T^{\rm ct}
\]
then the mean and noise are not obtained by keeping only the separately
computed component variances.  With \(i,j\) running over the displayed
components,
\begin{equation}
  j_a^{\rm full}
  =
  \sum_i
  \omega\!\left(T^i(\tau_a)\right),
  \qquad
  \mathsf N_{ab}^{\rm full}
  =
  \frac12
  \sum_{i,j}
  \omega\!\left(
    \left\{
      \Delta T^i(\tau_a),
      \Delta T^j(\tau_b)
    \right\}
  \right)_{\rm ren}.
  \label{eq:semiclassical-full-noise-component-sum}
\end{equation}
The off-diagonal \(i\ne j\) entries are part of the physical covariance.  They
are the fluctuation analogue of the missing Bogoliubov/contact terms in the
mean source equation, and they are required before the Ward test can be applied
to the full noise:
\begin{equation}
  \mathsf B_T j^{\rm full}=0,
  \qquad
  \mathsf B_T\mathsf N^{\rm full}=0,
  \qquad
  \mathsf N^{\rm full}\mathsf B_T^*=0 .
  \label{eq:semiclassical-full-package-ward-conditions}
\end{equation}
A projected partial covariance
\(\mathsf P_{\ker\mathsf B_T}\mathsf N^{\rm pot}\mathsf P_{\ker\mathsf B_T}^*\)
is therefore only a retained diagnostic or an explicit truncated model.  It is
not the fluctuation input of the interacting semiclassical equation unless the
omitted component covariances and cross-covariances have been proved to vanish
or have been included as controlled residuals.

Finite renormalization makes the distinction sharper.  A c-number local
curvature ambiguity
\[
  T_{\mu\nu}\mapsto T_{\mu\nu}+A_{\mu\nu}[g]\mathbf 1
\]
shifts \(j^{\rm full}\) and the gravitational EFT coordinates but drops out of
the connected noise.  By contrast, a finite composite-operator mixing
\[
  T_{\mu\nu}\mapsto
  T_{\mu\nu}+\sum_r c_r\,\mathcal O^{(r)}_{\mu\nu}
\]
changes the noise by cross terms:
\begin{equation}
\begin{aligned}
  \mathsf N'_{ab}
  &=
  \mathsf N_{ab}
  +
  \sum_r c_r
  \mathsf N_{ab}(T,\mathcal O^{(r)})
  +
  \sum_r c_r
  \mathsf N_{ab}(\mathcal O^{(r)},T) \\
  &\qquad
  +
  \sum_{r,s}c_rc_s\,
  \mathsf N_{ab}(\mathcal O^{(r)},\mathcal O^{(s)}).
\end{aligned}
  \label{eq:semiclassical-noise-finite-renormalization}
\end{equation}
Thus choosing the finite stress-tensor scheme is not only a mean-source
choice.  It also fixes which composite fluctuations and cross-correlations are
fed into the Einstein--Langevin covariance.

On a retained frequency window, the interacting semiclassical input is the
pair
\[
  h^{\rm full}(\omega)
  =
  \mathsf D(\omega+\ii0)^{-1}j^{\rm full}(\omega),
  \qquad
  \mathsf C_h^{\rm full}(\omega)
  =
  \mathsf D(\omega+\ii0)^{-1}
  \mathsf N^{\rm full}(\omega)
  \mathsf D(\omega+\ii0)^{-*},
\]
with \(\mathsf D\) built from the same gravitational EFT coordinates and
retarded response \(\mathsf K^{\rm ret}\).  A calculation has reached the
interacting backreaction level only when it supplies
\eqref{eq:semiclassical-full-interacting-package} in one scheme, or supplies
residual bounds for the missing state, product-extension, Ward, response, and
finite-renormalization data.
\end{controlledapproximation}

\begin{controlledapproximation}[Closed-time-path consistency of the interacting
backreaction package]
\label{ca:semiclassical-interacting-influence-package}
The three entries in
\eqref{eq:semiclassical-full-interacting-package} should be viewed as
derivatives of one renormalized closed-time-path influence functional, not as
three independently fitted matrices.  On the retained sector write
\[
  h_c=\frac12(h_++h_-),
  \qquad
  h_\Delta=h_+-h_- .
\]
After the state, product extension, local counterterms, and Ward reduction have
all been fixed in the same prescription, the quadratic influence functional has
the retained form
\begin{equation}
\begin{aligned}
  \Gamma_{\rm IF}^{(2)}
  &=
  \int h_\Delta^a j_a^{\rm full}
  +
  \int h_\Delta^a\,
       \mathsf R^{\rm ret}_{ab}\,
       h_c^b
  +
  \frac{\ii}{2}
  \int h_\Delta^a\,
       \mathsf N_{ab}^{\rm full}\,
       h_\Delta^b
  +R_{\rm IF}.
\end{aligned}
  \label{eq:semiclassical-ctp-influence-package}
\end{equation}
Here \(\mathsf R^{\rm ret}\) is shorthand, after choosing the retained
metric/stress pairing, for the typed interacting stress-response map
\(\mathsf R^{\rm ret}_{Th}:V_h\to V_T\) used in
\(\mathsf D_{\rm full}=\mathsf D_0-\mathsf R^{\rm ret}_{Th}\).  It includes
the declared local metric-variation terms, and \(R_{\rm IF}\) is the residual
from omitted products, higher cumulants, higher metric powers, and truncation
outside the retained window.

The form \eqref{eq:semiclassical-ctp-influence-package} encodes four physical
requirements.  First, normalization of the closed time path gives
\(\Gamma_{\rm IF}[h,h]=0\); hence there is no standalone \(h_c h_c\) kernel
and no mean term linear in \(h_c\).  Second, causality puts
\(\mathsf R^{\rm ret}_{ab}(t,t')=0\) unless the support of \(\tau_a\) lies in
the causal future of the support of \(\tau_b\).  Third, the imaginary quadratic
part is a covariance:
\begin{equation}
  v^a\mathsf N_{ab}^{\rm full}v^b\ge0
  \qquad
  \text{for every real retained test vector }v .
  \label{eq:semiclassical-ctp-noise-positivity}
\end{equation}
Fourth, diffeomorphism covariance acts on typed maps, not on one universal
matrix.  Let \(V_h\) be retained metric perturbations, \(V_T\) retained stress
coordinates, \(V_{\nabla T}\) retained divergence tests, and \(V_X\) retained
vector-field generators.  The stress-divergence map is
\(\mathsf B_T:V_T\to V_{\nabla T}\), the metric gauge generator is
\(\mathsf G_h:V_X\to V_h\), and the retained stress response is
\(\mathsf R^{\rm ret}_{Th}:V_h\to V_T\).  The finite version of
\eqref{eq:semiclassical-linearized-stress-ward} is
\begin{equation}
  \mathsf B_T j^{\rm full}=0,\qquad
  \mathsf B_T\mathsf N^{\rm full}=0,\qquad
  \mathsf N^{\rm full}\mathsf B_T^*=0,\qquad
  \mathsf B_T\mathsf R^{\rm ret}_{Th}
  +
  \mathsf Q_T=0,
  \label{eq:semiclassical-ctp-package-ward}
\end{equation}
where \(\mathsf Q_T:V_h\to V_{\nabla T}\) is the retained one-point/contact map
coming from \(\mathcal C_\nu[h;\bar T]\), from derivatives of the retarded
step function, and from equal-time stress commutators in the CTP product
extension.  A pure metric gauge perturbation obeys
\begin{equation}
  \mathsf R^{\rm ret}_{Th}\mathsf G_h=\mathsf L_T,
  \qquad
  \mathsf D_{\rm sc}\mathsf G_h=0,
  \qquad
  \mathsf D_{\rm sc}:=\mathsf D_g-\mathsf R^{\rm ret}_{Th},
  \label{eq:semiclassical-typed-gauge-response}
\end{equation}
where \(\mathsf L_T\) is the retained Lie derivative of the background
one-point stress tensor and \(\mathsf D_g\) is the gravitational/local-contact
metric Hessian in the same equation space.  Thus
\(\mathsf R^{\rm ret}_{Th}\mathsf G_h\) is generally not zero when
\(\bar T\ne0\); the gauge null vector is a null vector of the complete
semiclassical Hessian on a mean-equation background.

If a calculation wants ordinary reduced matrices, it must construct the
quotient.  Choose a metric reduction \(q_h:V_h\to V_h^{\rm red}\) with
\(\ker q_h\supset\operatorname{im}\mathsf G_h\), a section
\(s_h:V_h^{\rm red}\to V_h\), and a stress projection
\(p_T:V_T\to V_T^{\rm red}\) that kills the retained Lie-stress directions
\(p_T\mathsf L_T=0\) and is used only after the local/contact term
\(\mathsf Q_T\) has been included.  Then
\begin{equation}
  \mathsf R_{\rm red}^{\rm ret}
  :=
  p_T\mathsf R^{\rm ret}_{Th}s_h
  \label{eq:semiclassical-reduced-response-definition}
\end{equation}
is independent of the representative of \(s_h\) precisely because
\eqref{eq:semiclassical-typed-gauge-response} holds.  Only in such a
constructed quotient, or in a special background where
\(\mathsf Q_T=0=\mathsf L_T\) on the retained tests, does the tempting display
\(\mathsf B_T\mathsf R^{\rm ret}_{Th}=0\) become a valid reduced statement.  In
a stationary KMS state,
the same package must also satisfy the fluctuation--dissipation relation
\eqref{eq:semiclassical-fdt} for the interacting stress tensor in the chosen
scheme.  Thus a positive noise matrix and a causal retarded matrix are not yet
enough: their spectral parts must be compatible with the same state.

Once \eqref{eq:semiclassical-ctp-noise-positivity} holds, the
Einstein--Langevin source is the Hubbard--Stratonovich representative of the
imaginary part of \(\Gamma_{\rm IF}^{(2)}\):
\[
  \langle\xi_a\xi_b\rangle=\mathsf N^{\rm full}_{ab}.
\]
The stochastic equation is therefore not an additional assumption placed on top
of the interacting stress calculation.  It is the real-noise representation of
the same influence functional whose real part supplies the mean source and the
retarded dissipative kernel.  Any calculation that supplies \(j^{\rm full}\),
\(\mathsf N^{\rm full}\), and \(\mathsf R^{\rm ret}\) from different schemes,
or supplies a Ward-clean noise with an unrelated retarded kernel, has not yet
defined an interacting semiclassical backreaction datum.
\end{controlledapproximation}

\begin{controlledapproximation}[Causal backreaction small-gain control]
\label{ca:semiclassical-backreaction-small-gain-stability}
The interacting package above is still only source data until the metric
operator with stress-tensor feedback is controlled.  Work on one retained
Ward-clean frequency window \(I\).  Let \(\mathsf D_0(\omega)\) denote the
gauge-fixed gravitational and local-contact operator after constraint
reduction and reduction of order, before the nonlocal interacting
stress-tensor response is fed back.  Let
\(\mathsf R^{\rm ret}(\omega)\) denote the corresponding retained retarded
stress response from the same package
\eqref{eq:semiclassical-full-interacting-package}, with the factor \(1/2\)
and the sign convention of
\eqref{eq:semiclassical-retained-response-matrix} absorbed into
\(\mathsf R^{\rm ret}\).  Thus
\begin{equation}
  \mathsf D_{\rm full}(\omega)
  =
  \mathsf D_0(\omega)-\mathsf R^{\rm ret}(\omega).
  \label{eq:semiclassical-full-feedback-operator}
\end{equation}
The boundary-value estimate
\[
  \eta_{\partial I}
  =
  \sup_{\omega\in I}
  \left\lVert
    \mathsf D_0(\omega+\ii0)^{-1}
    \mathsf R^{\rm ret}(\omega+\ii0)
  \right\rVert_2
\]
is only a real-window diagnostic.  It can prove pointwise invertibility of
sampled matrices on \(I\), but it does not by itself prove a causal inverse,
exclude an upper-half-plane pole, or control the homogeneous initial-value
problem.

For a retained \emph{retarded} inverse, impose the small-gain estimate before
taking the boundary value.  Let
\(\Omega_+=\{z\in\mathbb C:\operatorname{Im}z>0\}\).  Assume that
\(\mathsf D_0(z)^{-1}\) and \(\mathsf R^{\rm ret}(z)\) are operator-valued
Hardy functions on \(\Omega_+\) in the reduced sector, with tempered boundary
values, retarded inverse-Fourier/Laplace transforms, and no
upper-half-plane poles after the cutoff-scale roots have been removed by the
EFT reduction-of-order prescription.  Assume
\begin{equation}
  M_{0,+}
  =
  \sup_{z\in\Omega_+}
  \left\lVert
    \mathsf D_0(z)^{-1}
  \right\rVert_2
  <\infty,
  \qquad
  \eta_+
  =
  \sup_{z\in\Omega_+}
  \left\lVert
    \mathsf D_0(z)^{-1}
    \mathsf R^{\rm ret}(z)
  \right\rVert_2
  <1 .
  \label{eq:semiclassical-small-gain-hypothesis}
\end{equation}
Then the full retained inverse exists as an upper-half-plane analytic Hardy
function,
\begin{equation}
\begin{aligned}
  \mathsf D_{\rm full}(z)^{-1}
  &=
  \left(
    \mathbf 1-\mathsf D_0(z)^{-1}\mathsf R^{\rm ret}(z)
  \right)^{-1}
  \mathsf D_0(z)^{-1}
  \\
  &=
  \sum_{m\ge0}
  \left(\mathsf D_0(z)^{-1}\mathsf R^{\rm ret}(z)\right)^m
  \mathsf D_0(z)^{-1},
  \\
  \left\lVert\mathsf D_{\rm full}(z)^{-1}\right\rVert_2
  &\le
  \frac{M_{0,+}}{1-\eta_+}.
\end{aligned}
  \label{eq:semiclassical-small-gain-inverse}
\end{equation}
The Neumann series converges normally in \(\Omega_+\), so it cannot generate an
upper-half-plane pole.  The Hardy boundary-value theorem, equivalently the
Paley--Wiener support theorem for the inverse Fourier--Laplace transform,
therefore makes the boundary kernel retarded.  Real-frequency singular-value
samples do not supply this support statement.

If the analytic estimate is proved only in a tube above \(I\), the conclusion is
only a windowed forced-response statement unless one separately excludes
upper-half-plane poles elsewhere in the reduced sector and bounds the coupling
to omitted modes.  A full initial-value stability statement also needs a
homogeneous semigroup or energy estimate for the retained and omitted
gravitational variables.  The small-gain bound below controls the retarded
forced response and the stochastic pushforward in the declared sector; it is
not, by itself, a global stability theorem for the full semiclassical PDE.

For a Ward-clean interacting source and noise package,
\[
  \mathsf B_T j^{\rm full}=0,\qquad
  \mathsf B_T\mathsf N^{\rm full}=0,\qquad
  \mathsf N^{\rm full}\mathsf B_T^*=0,
\]
the retained mean response and metric covariance obey
\begin{equation}
  h^{\rm full}
  =
  \mathsf D_{\rm full}^{-1}j^{\rm full},
  \qquad
  \mathsf C_h^{\rm full}
  =
  \mathsf D_{\rm full}^{-1}
  \mathsf N^{\rm full}
  \mathsf D_{\rm full}^{-*},
  \label{eq:semiclassical-small-gain-response-covariance}
\end{equation}
with
\begin{equation}
  \lVert h^{\rm full}\rVert_2
  \le
  \frac{M_{0,+}}{1-\eta_+}\lVert j^{\rm full}\rVert_2,
  \qquad
  \operatorname{tr}\mathsf C_h^{\rm full}
  \le
  \frac{M_{0,+}^2}{(1-\eta_+)^2}
  \operatorname{tr}\mathsf N^{\rm full}.
  \label{eq:semiclassical-small-gain-mean-noise-bound}
\end{equation}
If a truncated calculation supplies only
\(\mathsf N^{\rm ret}\) and a positive trace residual
\(\operatorname{tr}\mathsf N^{\rm miss}\le r_N\), the last bound is replaced
by the same expression with
\(\operatorname{tr}\mathsf N^{\rm ret}+r_N\).  This is the place where a
finite interacting example becomes a backreaction statement: the retarded
feedback must be small enough in the causal norm, or otherwise explicitly
invertible with an independently proved retarded inverse, in the same
Ward-clean sector in which the mean source and connected noise are defined.
If the projection discards modes, their source and noise leakage must be
included as residual budgets before the retained covariance is interpreted as a
physical metric-fluctuation bound.

Failure of any part of this test has physical content.  If
\(\eta_+\ge1\) and no sharper causal inverse estimate is supplied, the
perturbative feedback expansion is not controlled.  If
\(\mathsf D_{\rm full}^{-1}\) has an upper-half-plane pole, the retained
background is linearly unstable in the semiclassical problem even if every
sampled real-frequency matrix in \(I\) is invertible.  If
\(\operatorname{tr}\mathsf C_h^{\rm full}\) is comparable to the square of
the metric reference scale, or the induced curvature covariance is comparable
to the square of the background curvature scale, the mean semiclassical
metric is not a self-averaging output of the declared state.  Ward projection
alone cannot diagnose these failures: a projected source may be conserved
while the retarded feedback is unstable, and a conserved connected noise
kernel may still be amplified beyond the validity of the linearized metric
chart.
\end{controlledapproximation}

\begin{controlledapproximation}[Nonlinear finite-window backreaction chart]
\label{ca:semiclassical-nonlinear-backreaction-chart}
The small-gain test controls the linearized operator.  A semiclassical
solution also needs a chart in which the nonlinear gravitational terms, the
metric-dependence of the interacting state, and the omitted retained
stress-tensor residuals stay inside the same response window.  Keep the
Ward-clean retained sector and write
\[
  \mathsf G_{\rm full}
  =
  \mathsf D_{\rm full}^{-1},
  \qquad
  M_{\rm full}
  =
  \frac{M_{0,+}}{1-\eta_+},
  \qquad
  h_{\rm lin}
  =
  \mathsf G_{\rm full}j^{\rm full}.
\]
After gauge fixing, constraint reduction, and reduction of order, suppose the
retained nonlinear equation in a ball
\(B_h(r_h)=\{h:\lVert h\rVert\le r_h\}\) has the form
\begin{equation}
  h=\Phi(h)
  =
  h_{\rm lin}
  +
  \mathsf G_{\rm full}
  \left(
    \mathcal Q_2(h,h)+\mathcal S_{\rm st}(h)+\mathcal R_{\rm nl}(h)
  \right),
  \label{eq:semiclassical-nonlinear-fixed-map}
\end{equation}
where \(\mathcal Q_2\) contains the retained quadratic metric terms and
second metric variations of the source, \(\mathcal S_{\rm st}\) is the
linearized state-transport contribution in the chosen Hadamard chart, and
\(\mathcal R_{\rm nl}\) is the declared residual from omitted higher retained
orders and unresolved stress-tensor products.  All three terms are required to
land in the Ward-clean sector; otherwise the fixed-point problem has left the
same constraint-reduced space in which
\(\mathsf G_{\rm full}\) was estimated.

Assume that, for \(h,\tilde h\in B_h(r_h)\),
\begin{equation}
\begin{gathered}
  \lVert\mathcal Q_2(h,h)\rVert\le C_2\lVert h\rVert^2,
  \qquad
  \lVert
    \mathcal Q_2(h,h)-\mathcal Q_2(\tilde h,\tilde h)
  \rVert
  \le
  2C_2r_h\lVert h-\tilde h\rVert,
  \\
  \lVert\mathcal S_{\rm st}(h)\rVert\le L_{\rm st}\lVert h\rVert,
  \qquad
  \lVert
    \mathcal S_{\rm st}(h)-\mathcal S_{\rm st}(\tilde h)
  \rVert
  \le
  L_{\rm st}\lVert h-\tilde h\rVert,
  \qquad
  \sup_{B_h(r_h)}\lVert\mathcal R_{\rm nl}\rVert\le\epsilon_{\rm nl},
  \\
  \lVert
    \mathcal R_{\rm nl}(h)-\mathcal R_{\rm nl}(\tilde h)
  \rVert
  \le
  L_{\rm R}\lVert h-\tilde h\rVert .
\end{gathered}
  \label{eq:semiclassical-nonlinear-chart-bounds}
\end{equation}
Then the retained equation is controlled on this finite window when
\begin{equation}
  \lVert h_{\rm lin}\rVert
  +
  M_{\rm full}
  \left(C_2r_h^2+L_{\rm st}r_h+\epsilon_{\rm nl}\right)
  \le r_h,
  \qquad
  \kappa_{\rm nl}
  =
  M_{\rm full}(2C_2r_h+L_{\rm st}+L_{\rm R})<1 .
  \label{eq:semiclassical-nonlinear-contraction-test}
\end{equation}
The first inequality says that the declared source and residuals stay inside
the metric chart; the second says that nonlinear feedback and state transport
cannot undo the linear small-gain estimate once the omitted retained residual
has also been controlled as a function of the metric perturbation.  The last
Lipschitz line is essential: a bounded \(h\)-dependent residual gives only a
self-map estimate, not a contraction.  If the residual is an additive
\(h\)-independent defect, this is the special case \(L_{\rm R}=0\); if only a
size bound is available, the finite chart must not claim Picard uniqueness
without replacing the argument by separate compactness and continuity
hypotheses.  Under the displayed contraction hypothesis the Picard sequence
defined by \(\Phi\) converges to a unique retained solution in \(B_h(r_h)\), and
the nonlinear correction to the linear response is bounded by
\begin{equation}
  \lVert h-h_{\rm lin}\rVert
  \le
  \frac{
    M_{\rm full}\left(C_2r_h^2+L_{\rm st}r_h+\epsilon_{\rm nl}\right)
  }{1-\kappa_{\rm nl}} .
  \label{eq:semiclassical-nonlinear-correction-bound}
\end{equation}

The stochastic part needs the same bookkeeping.  If \(r_N\) is the trace
budget for omitted connected stress noise and \(\Delta_N(r_h)\) is the
declared nonlinear/state-transport correction to the noise in this chart, the
metric-fluctuation validity test is refined to
\begin{equation}
  \operatorname{tr}\mathsf C_h^{\rm full}
  +
  \frac{M_{\rm full}^2}{(1-\kappa_{\rm nl})^2}\,r_N
  +
  \Delta_N(r_h)
  \ll r_h^2 .
  \label{eq:semiclassical-nonlinear-noise-validity}
\end{equation}
Thus a finite backreaction statement consists of a matched package: a
scheme-fixed interacting source/noise/response calculation, a Ward-clean
constraint reduction, a bounded feedback inverse, a state-transport radius,
and residual budgets small enough for both the mean solution and its metric
fluctuations to remain in the declared semiclassical chart.  A full
semiclassical existence theorem would replace these finite retained
inequalities by topology, hyperbolic evolution, and Hadamard-state
propagation estimates for the infinite-dimensional problem.
\end{controlledapproximation}

\begin{controlledapproximation}[Retained metric-observable output]
\label{ca:semiclassical-retained-metric-observable-output}
A controlled backreaction calculation should end with an observable, not only
with a small metric perturbation.  Let \(X\) be a gauge-invariant retained
metric observable, or a gauge-fixed observable after the same constraint
reduction used to define \(\mathsf G_{\rm full}\).  Examples are a smeared
curvature scalar, a quasi-local mass coordinate, a horizon-area functional in
the chosen perturbative chart, or the FLRW Hubble coordinate tested against
the reduced Friedmann constraint.  In the retained chart write
\begin{equation}
  X[g_0+h]
  =
  X[g_0]
  +
  \ell_X^a h_a
  +
  \frac12 Q^X_{ab}h^a h^b
  +
  R_X(h),
  \qquad
  |R_X(h)|\le\epsilon_X
  \quad(h\in B_h(r_h)).
  \label{eq:semiclassical-observable-chart}
\end{equation}
The linear functional \(\ell_X\), the quadratic form \(Q^X\), and the
remainder bound are part of the observable definition in the same metric
chart as the backreaction solution.  A coordinate component of \(h\) that has
not survived the gauge and constraint reduction is not an observable output of
the calculation.

The stochastic average needs one more datum: the probability law whose
expectation is being taken.  In the retained finite vector space write
\begin{equation}
  H=h_*+Z,
  \qquad
  Z\sim\mathcal N(0,\mathsf C_h),
  \qquad
  \rho_h=r_h-\lVert h_*\rVert>0,
  \qquad
  \mathcal E_h=\{H\notin B_h(r_h)\}.
  \label{eq:semiclassical-observable-gaussian-law}
\end{equation}
Then \(\mathcal E_h\subset\{\lVert Z\rVert>\rho_h\}\), so the local Taylor
bound in \eqref{eq:semiclassical-observable-chart} controls an unstopped
expectation only after the chart-exit probability and the out-of-chart
observable growth have also been supplied.  At the minimum,
\begin{equation}
  \mathbb P(\mathcal E_h)
  \le
  \frac{\operatorname{tr}\mathsf C_h}{\rho_h^2},
  \label{eq:semiclassical-observable-markov-exit-bound}
\end{equation}
and, for the Euclidean retained norm, the sharper Gaussian tail
\begin{equation}
  \mathbb P\!\left(
    \lVert Z\rVert^2
    \ge
    \operatorname{tr}\mathsf C_h
    +2\sqrt{\lVert\mathsf C_h\rVert_{\rm HS}^2s}
    +2\lVert\mathsf C_h\rVert_{\rm op}s
  \right)
  \le e^{-s}
  \qquad(s>0)
  \label{eq:semiclassical-observable-gaussian-exit-bound}
\end{equation}
can be used.  A valid observable statement must choose one of three exits from
this problem: a global integrability bound on the out-of-chart remainder, a
stopped or truncated Einstein--Langevin process with an explicit stopping
error, or a conditional-on-chart prediction together with the displayed exit
probability.

Let \(h_*\) be the retained mean solution supplied by
Controlled Approximation~\ref{ca:semiclassical-nonlinear-backreaction-chart}
and let \(\mathsf C_h\) be the corresponding Einstein--Langevin metric
covariance, including the missing-noise and nonlinear/state-transport budgets
of \eqref{eq:semiclassical-nonlinear-noise-validity}.  To the same retained
order, the mean observable shift is
\begin{equation}
\begin{aligned}
  \Delta_X^{\rm mean}
  &=
  \ell_X^a h_{*,a}
  +
  \frac12 Q^X_{ab}h_*^a h_*^b
  +
  \frac12\operatorname{tr}(Q^X\mathsf C_h)
  +
  R_X^{\rm mean},
  \\
  |R_X^{\rm mean}|
  &\le
  \epsilon_X
  +
  \frac12\lVert Q^X\rVert\,r_N^{(h)}
  +
  r_X^{\rm higher}
  +
  r_X^{\rm exit}.
\end{aligned}
  \label{eq:semiclassical-observable-mean-shift}
\end{equation}
Here \(r_N^{(h)}\) is the retained metric-covariance trace budget not already
included in the displayed \(\mathsf C_h\), \(r_X^{\rm higher}\) covers the
discarded higher cumulants and higher metric powers inside the chart, and
\(r_X^{\rm exit}\) bounds the out-of-chart contribution.  For a stopped
process with \(|X|\le B_X\) on the compared stopped/unstopped laws one may take
\(r_X^{\rm exit}\le 2B_X\mathbb P(\mathcal E_h)\).  For a global growth bound
\(|R_X^{\rm out}(h)|\le G_X(h)\), one must instead supply
\(\mathbb E[G_X(H)\mathbf1_{\mathcal E_h}]\).  Without one of these
quantities the local number \(\epsilon_X\) is not an expectation bound,
because a nondegenerate Gaussian has support outside \(B_h(r_h)\).  The term
\(\frac12\operatorname{tr}(Q^X\mathsf C_h)\) is the fluctuation bias of the
metric observable.  It is absent in a deterministic mean-only treatment, but
it is a physical contribution whenever the Einstein--Langevin fluctuations
are kept at the same order as the quadratic observable response.

The leading observable covariance between two retained observables \(X,Y\) is
\begin{equation}
  \operatorname{Cov}(X,Y)
  =
  \ell_X^a(\mathsf C_h)_{ab}\ell_Y^b
  +
  R_{XY}^{\rm cov},
  \qquad
  |R_{XY}^{\rm cov}|
  \le r_{XY}^{\rm quad}+r_{XY}^{\rm higher}+r_{XY}^{\rm exit}.
  \label{eq:semiclassical-observable-covariance}
\end{equation}
The exit term is again part of the covariance budget, not an optional
afterthought: a rare out-of-chart metric perturbation can dominate a nonlinear
observable even when \(\operatorname{tr}\mathsf C_h\) is small.

An evaporation rate, a cosmological expansion shift, or a Casimir
backreaction correction is a self-averaging or directly detectable mean shift
only after the observable signal is compared with its induced stress-noise
uncertainty:
\begin{equation}
  |\Delta_X^{\rm mean}|
  \gg
  \sqrt{
    \ell_X^a(\mathsf C_h)_{ab}\ell_X^b
    +r_{XX}^{\rm quad}
    +r_{XX}^{\rm tail}
  } .
  \label{eq:semiclassical-observable-signal-to-noise}
\end{equation}
This inequality is not the validity criterion for the stochastic prediction.
If the law, chart-exit, and integrability budgets above are controlled but the
inequality fails, the calculation still predicts a probability distribution
for \(X\); it is simply noise dominated, not self-averaging and not a clean
measurement of the mean shift.  If the exit or integrability budget is missing,
by contrast, the displayed expectation and covariance are not controlled at
all.  If the noise used in \(\mathsf C_h\) is only a projected component
covariance rather than the scheme-fixed interacting stress-tensor noise of
\eqref{eq:semiclassical-full-interacting-package}, neither the probability law
nor the signal-to-noise test is the physical one.
\end{controlledapproximation}

\section{Validity Conditions And Reduction Of Order}

The semiclassical approximation requires more than small curvature.  A
controlled statement includes
\begin{equation}
\begin{gathered}
  \ell_{\rm curv}\gg \ell_{\rm UV},\\
  |\langle T_{\mu\nu}^{\rm ren}\rangle|
  \text{ small compared with the gravitational EFT cutoff scales},\\
  N(f,f)^{1/2}
  \text{ small for the smeared observables used to test the metric},\\
  \text{observable chart-exit and tail budgets controlled},\\
  \text{signal-to-noise controlled only when claiming a resolved mean shift},\\
  \text{state stability under the metric perturbations considered},\\
  \text{hyperbolic well-posedness after higher-curvature terms are treated}.
\end{gathered}
  \label{eq:semiclassical-validity-conditions}
\end{equation}
Curvature-squared terms contain four derivatives of the metric.  Treated as
exact equations, they introduce additional modes at the cutoff scale.  In a
low-energy gravitational EFT these modes are not new physical initial data.
They are removed by reduction of order: solve the leading second-order
equation, use it to replace higher derivatives in the curvature-squared
terms, and keep only the terms through the fixed EFT order.  Equivalently,
work perturbatively in \(\alpha\) and \(\beta\) and discard solutions that
are singular as \(\alpha,\beta\to0\).

\section{Backreaction Specification}

A backreaction problem is specified by
\[
  (M,g,\omega,T_{\mu\nu}^{\rm ren},
  G_N,\Lambda,\alpha,\beta,\mathcal B),
\]
where \(\mathcal B\) denotes boundary or initial data, gauge conditions, and
state-transport prescriptions.  The output is a solution space with
stability, state-dependence, and renormalization-coordinate dependence.
Black-hole evaporation, cosmological particle creation, and Casimir
backreaction are applications of this datum.

\begin{openproblem}[Backreaction theorem]
\label{op:backreaction-theorem}
Formulate and prove a semiclassical backreaction theorem for an interacting
Hadamard QFT with renormalized stress tensor, controlled stress-tensor
fluctuations, local covariance, and a hyperbolic gravitational EFT evolution.
The theorem should specify the topology on metric perturbations and states,
the allowed renormalization coordinates, the treatment of higher-derivative
terms, and the sense in which the stochastic approximation captures the
leading connected stress-tensor fluctuations.
\end{openproblem}
