\chapter{Trace Anomalies and Background Variations}
\label{ch:curved-trace-anomalies}

\ConventionsEuclidean

The preceding chapter defined renormalized stress tensors on curved
backgrounds by locally covariant point splitting.  The present chapter studies
the trace of that stress tensor.  The trace anomaly is a local response of
the renormalized source functional to a Weyl variation of the background
metric.  It is part of the same curved-background QFT datum as the stress
tensor, its contact terms, and its finite curvature counterterms.
The upstream inputs are therefore the point-split stress tensor,
conservation/locality of the chosen finite counterterms, and the local
Hadamard subtraction class.  This chapter changes the background metric and
tracks the local Weyl response; it does not replace the microlocal state
condition or the pAQFT construction of interacting composite fields.

\section{Metric Source Convention}

Let \(M\) be a closed smooth \(D\)-manifold and let \(g\) be a Riemannian
metric.  Write
\[
  \mathcal W[g,\lambda]
\]
for the renormalized Euclidean effective action, including all declared
background sources \(\lambda\) besides the metric.  In a path-integral
regularization this is \(-\log Z[g,\lambda]\), after local counterterms have
been fixed.  The stress-tensor expectation value is defined by
\begin{equation}
  \delta_g\mathcal W[g,\lambda]
  =
  -\frac12\int_M d^D x\,\sqrt g\,
  \langle T^{\mu\nu}(x)\rangle_{g,\lambda}\,
  \delta g_{\mu\nu}(x).
  \label{eq:curved-trace-stress-convention}
\end{equation}
Equivalently,
\[
  \langle T^{\mu\nu}(x)\rangle_{g,\lambda}
  =
  -\frac{2}{\sqrt g}
  \frac{\delta\mathcal W}{\delta g_{\mu\nu}(x)} .
\]
For a local Weyl parameter \(\sigma\in C^\infty(M)\),
\[
  \delta_\sigma g_{\mu\nu}=2\sigma g_{\mu\nu}.
\]
Substituting this variation into
\eqref{eq:curved-trace-stress-convention} gives
\begin{equation}
  \delta_\sigma\mathcal W[g,\lambda]
  =
  -\int_M d^D x\,\sqrt g\,
  \sigma(x)\,\langle T^\mu{}_\mu(x)\rangle_{g,\lambda}.
  \label{eq:curved-trace-weyl-response}
\end{equation}

\begin{definition}[Trace-anomaly density]
\label{def:curved-trace-anomaly-density}
For a theory whose classical curved-background action is Weyl invariant
after the assigned transformations of all sources \(\lambda\), the
renormalized trace-anomaly density \(\mathcal A[g,\lambda]\) is the local
scalar density coefficient defined by
\[
  \langle T^\mu{}_\mu(x)\rangle_{g,\lambda}
  =
  \mathcal A[g,\lambda](x)
\]
in the vacuum or state-independent part of the Weyl response.  Thus
\begin{equation}
  \delta_\sigma\mathcal W[g,\lambda]
  =
  -\int_M d^D x\,\sqrt g\,\sigma\,\mathcal A[g,\lambda].
  \label{eq:curved-trace-anomaly-definition}
\end{equation}
\end{definition}

If the theory has mass parameters, relevant couplings, or fields assigned
nontrivial Weyl transformations, the trace Ward identity contains additional
operator terms such as \(\beta^i\mathcal O_i\), masses times composite
operators, and improvement terms.  This chapter isolates the purely local
curvature part which remains in a conformal theory or in the state-independent
part of a locally covariant stress tensor.

\section{Wess--Zumino Consistency}

Let
\[
  \mathbf A_\sigma[g,\lambda]
  :=
  \int_M d^D x\,\sqrt g\,\sigma\,\mathcal A[g,\lambda].
\]
The Weyl transformations form an abelian Lie algebra:
\[
  [\delta_{\sigma_1},\delta_{\sigma_2}]g_{\mu\nu}=0
\]
when the source transformations are also abelian.  Since
\(\mathcal W[g,\lambda]\) is a functional, the commutator of its two Weyl
variations must vanish.

\begin{remark}[Wess--Zumino consistency]
\label{rem:curved-trace-wz-consistency}
Assume the Weyl variation of \(\mathcal W\) is local and has the form
\eqref{eq:curved-trace-anomaly-definition}.  Then
\begin{equation}
  \delta_{\sigma_1}\mathbf A_{\sigma_2}
  -
  \delta_{\sigma_2}\mathbf A_{\sigma_1}
  =
  0.
  \label{eq:curved-trace-wz-consistency}
\end{equation}
If a finite local counterterm \(B[g,\lambda]\) is added to
\(\mathcal W\), then
\begin{equation}
  \mathbf A_\sigma\longmapsto
  \mathbf A_\sigma-\delta_\sigma B.
  \label{eq:curved-trace-counterterm-coboundary}
\end{equation}
Therefore the intrinsic trace anomaly is a class in local Weyl cohomology:
local solutions of \eqref{eq:curved-trace-wz-consistency} modulo local Weyl
variations of counterterms.
\end{remark}

The Weyl transformations are abelian on the metric-source space under the
present hypotheses:
\[
  [\delta_{\sigma_1},\delta_{\sigma_2}]=0 .
\]
Using \(\delta_\sigma\mathcal W=-\mathbf A_\sigma\),
\[
  0
  =
  [\delta_{\sigma_1},\delta_{\sigma_2}]\mathcal W
  =
  -\delta_{\sigma_1}\mathbf A_{\sigma_2}
  +\delta_{\sigma_2}\mathbf A_{\sigma_1}.
\]
This is \eqref{eq:curved-trace-wz-consistency}.  If
\(\mathcal W'=\mathcal W+B\), then
\[
  -\mathbf A'_\sigma
  =
  \delta_\sigma\mathcal W'
  =
  -\mathbf A_\sigma+\delta_\sigma B,
\]
which gives \eqref{eq:curved-trace-counterterm-coboundary}.

The consistency condition is a local cohomological statement only after
locality has been supplied.  In
perturbation theory, locality follows from the locality of counterterms.  For
free fields on curved backgrounds, it follows from the local Hadamard
parametrix or, equivalently, from local heat-kernel coefficients.  In a
nonperturbative locally covariant QFT, locality is an additional structural
property of the background dependence.

\section{Curvature Variations under Weyl Rescaling}

The classification in low dimensions uses a few explicit Weyl variations.
\paragraph{Weyl variation of the \(R^2\) counterterm.}
On a \(D\)-dimensional Riemannian manifold,
\begin{align}
  \delta_\sigma\sqrt g
  &=
  D\sigma\sqrt g,
  \label{eq:weyl-variation-volume}\\
  \delta_\sigma R
  &=
  -2\sigma R-2(D-1)\nabla^2\sigma,
  \label{eq:weyl-variation-ricci-scalar}
\end{align}
where \(\nabla^2=\nabla_\mu\nabla^\mu\).  In \(D=4\), these imply
\begin{equation}
  \delta_\sigma\int_M d^4x\,\sqrt g\,R^2
  =
  -12\int_M d^4x\,\sqrt g\,\sigma\,\nabla^2R
  \label{eq:r-squared-counterterm-weyl-variation}
\end{equation}
on a closed manifold.
The variation of the volume form follows from
\(\delta\sqrt g=\frac12\sqrt g\,g^{\mu\nu}\delta g_{\mu\nu}\).  The standard
variation of the scalar curvature under a metric variation \(h_{\mu\nu}\) is
\[
  \delta R
  =
  -R^{\mu\nu}h_{\mu\nu}
  +\nabla^\mu\nabla^\nu h_{\mu\nu}
  -\nabla^2(g^{\mu\nu}h_{\mu\nu}).
\]
Putting \(h_{\mu\nu}=2\sigma g_{\mu\nu}\) gives
\eqref{eq:weyl-variation-ricci-scalar}.  In \(D=4\),
\[
  \delta_\sigma(\sqrt g\,R^2)
  =
  4\sigma\sqrt g\,R^2
  +2\sqrt g\,R(-2\sigma R-6\nabla^2\sigma)
  =
  -12\sqrt g\,R\nabla^2\sigma .
\]
Integrating by parts twice gives
\eqref{eq:r-squared-counterterm-weyl-variation}.

Equation~\eqref{eq:r-squared-counterterm-weyl-variation} is the concrete
reason why the coefficient of \(\nabla^2R\) in four dimensions is a scheme
coordinate.  Adding \(\kappa(16\pi^2)^{-1}\int\sqrt g\,R^2\) shifts the
coefficient \(b\) in the convention below by \(12\kappa\).

\section{Two-Dimensional Anomaly}

In two dimensions, the local parity-even scalar of Weyl weight two is the
Ricci scalar.  The trace anomaly of a unitary two-dimensional conformal field
theory is written
\begin{equation}
  \langle T^\mu{}_\mu\rangle_g
  =
  -\frac{c}{24\pi}R .
  \label{eq:curved-two-dimensional-trace-anomaly}
\end{equation}
The coefficient \(c\) is the Virasoro central charge when the stress tensor
is normalized by
\[
  \langle T(z)T(0)\rangle=\frac{c}{2z^4}
\]
on the flat complex plane.

Let \(\hat g_{\mu\nu}=e^{2\tau}g_{\mu\nu}\).  The functional
\[
  S_{\rm WZ}^{(2)}[\tau;g]
  =
  -\frac{c}{24\pi}
  \int_M d^2x\,\sqrt g\,\bigl(\tau R+(\nabla\tau)^2\bigr)
\]
satisfies
\[
  S_{\rm WZ}^{(2)}[\tau+\sigma;g]
  -
  S_{\rm WZ}^{(2)}[\tau;g]
  =
  -\frac{c}{24\pi}
  \int_M d^2x\,\sqrt{\hat g}\,\sigma\,R[\hat g]
  +O(\sigma^2)
\]
for infinitesimal \(\sigma\).  Hence it integrates the anomaly
\eqref{eq:curved-two-dimensional-trace-anomaly}.
In two dimensions,
\[
  \sqrt{\hat g}\,R[\hat g]
  =
  \sqrt g\,(R-2\nabla^2\tau).
\]
Varying \(S_{\rm WZ}^{(2)}\) with respect to \(\tau\) gives
\[
  \delta_\sigma S_{\rm WZ}^{(2)}
  =
  -\frac{c}{24\pi}
  \int_M\sqrt g\,\sigma(R-2\nabla^2\tau)
  =
  -\frac{c}{24\pi}
  \int_M\sqrt{\hat g}\,\sigma R[\hat g],
\]
after integrating by parts in the \((\nabla\tau)^2\) term.

\section{Four-Dimensional Parity-Even Anomaly}

In four dimensions define
\[
  E_4
  =
  R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma}
  -4R_{\mu\nu}R^{\mu\nu}
  +R^2
\]
and
\[
  W^2
  =
  W_{\mu\nu\rho\sigma}W^{\mu\nu\rho\sigma}
  =
  R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma}
  -2R_{\mu\nu}R^{\mu\nu}
  +\frac13R^2 .
\]
For a parity-even four-dimensional conformal theory on a closed manifold,
the metric trace anomaly has the form
\begin{equation}
  \langle T^\mu{}_\mu\rangle_g
  =
  \frac{1}{16\pi^2}
  \left(
  c\,W^2-a\,E_4+b\,\nabla^2R
  \right)
  \label{eq:curved-four-dimensional-trace-anomaly}
\end{equation}
plus terms involving other background fields when such sources are present.

\begin{remark}[Four-dimensional cohomology representatives]
\label{rem:curved-four-dimensional-weyl-classes}
Modulo local Weyl variations of dimension-four metric counterterms and total
derivatives, the parity-even metric anomaly on a closed four-manifold has the
two nontrivial representatives \(E_4\) and \(W^2\).  The term
\(\nabla^2R\) is shifted by the local counterterm \(\int\sqrt g\,R^2\).
\end{remark}

The dimension-four parity-even scalar curvature basis is spanned by
\[
  R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma},
  \qquad
  R_{\mu\nu}R^{\mu\nu},
  \qquad
  R^2,
  \qquad
  \nabla^2R .
\]
The two combinations \(E_4\) and \(W^2\) displayed above are linearly
independent.  The integral of \(E_4\) is topological on a closed
four-manifold, so its Weyl variation is a total derivative and satisfies
Wess--Zumino consistency.  The density \(\sqrt g\,W^2\) is pointwise Weyl
invariant in four dimensions, so it also satisfies consistency.  The
\(\nabla^2R\) term is the Weyl variation of the \(R^2\) counterterm by
\eqref{eq:r-squared-counterterm-weyl-variation}.  These three directions
span the basis modulo a choice of representative, leaving \(E_4\) and \(W^2\)
as the nontrivial local cohomology classes in this sector.

If background gauge fields are present, further type-B terms appear.  For a
dimensionless background gauge field \(A\), the density
\(\operatorname{tr}_{\mathcal R}F_{\mu\nu}F^{\mu\nu}\) has Weyl weight four,
so its coefficient is constrained by the flavor current two-point
normalization and by the choice of current-source convention.  Parity-odd
terms and manifolds with boundary require a separate orientation and boundary
classification.

\section{Heat-Kernel Derivation for Free Conformal Fields}

The following computation gives a controlled derivation of the free-field
entries in \eqref{eq:curved-four-dimensional-trace-anomaly}.  Let \(P\) be a
positive Laplace-type operator on a vector bundle \(E\to M\), written
\[
  P=-(g^{\mu\nu}\nabla_\mu\nabla_\nu+\mathcal E).
\]
Let \(\Omega_{\mu\nu}=[\nabla_\mu,\nabla_\nu]\).  On a closed four-manifold,
the local heat-kernel expansion is
\[
  {\rm Tr}\,e^{-sP}
  \sim
  \frac{1}{(4\pi s)^2}
  \sum_{n\ge0}s^n
  \int_M d^4x\,\sqrt g\,
  {\rm tr}_E\,a_{2n}(P).
\]
The coefficient needed for the logarithmic Weyl response is
\begin{align}
  {\rm tr}_E\,a_4(P)
  &=
  {\rm tr}_E\left[
  \frac{1}{360}
  \left(
  5R^2-2R_{\mu\nu}R^{\mu\nu}
  +2R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma}
  \right)\mathbf 1
  \right.
  \notag\\
  &\hspace{4em}\left.
  +\frac16R\mathcal E+\frac12\mathcal E^2
  +\frac1{12}\Omega_{\mu\nu}\Omega^{\mu\nu}
  \right]
  +\nabla_\mu J^\mu .
  \label{eq:curved-heat-kernel-a4}
\end{align}
The divergent logarithmic term in \(\frac12\log\det P\) is proportional to
\(\int{\rm tr}\,a_4(P)\).  For Grassmann fields and gauge fields one includes
the corresponding minus signs, square-root factors, and ghost determinants.

For a real conformally coupled scalar in four dimensions,
\[
  P_\phi=-\nabla^2+\frac16R
  =
  -(\nabla^2+\mathcal E),
  \qquad
  \mathcal E=-\frac16R,
\]
and \(\Omega_{\mu\nu}=0\).  The non-derivative part of
\({\rm tr}\,a_4(P_\phi)\) is
\[
  \frac{1}{180}
  \left(
  R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma}
  -
  R_{\mu\nu}R^{\mu\nu}
  \right).
\]
Equivalently,
\[
  a_{\rm scalar}=\frac1{360},
  \qquad
  c_{\rm scalar}=\frac1{120}.
\]
Substituting \(\mathcal E=-R/6\) and \(\Omega_{\mu\nu}=0\) into
\eqref{eq:curved-heat-kernel-a4} gives the \(R^2\) coefficient
\[
  \frac{5}{360}-\frac{1}{36}+\frac{1}{72}=0.
\]
The remaining terms are
\[
  \frac{2}{360}R_{\mu\nu\rho\sigma}^2
  -
  \frac{2}{360}R_{\mu\nu}^2
  =
  \frac1{180}(R_{\mu\nu\rho\sigma}^2-R_{\mu\nu}^2).
\]
Using
\[
  cW^2-aE_4
  =
  (c-a)R_{\mu\nu\rho\sigma}^2
  +(-2c+4a)R_{\mu\nu}^2
  +\left(\frac c3-a\right)R^2
\]
with \(a=1/360\), \(c=1/120\), gives the same curvature combination.

For the remaining free conformal fields, the same heat-kernel formula is
applied to the squared Dirac operator and to the gauge-fixed Maxwell complex.
The fermion sign and the vector ghost subtraction are part of the determinant
calculation:
\[
  \mathcal W_{\rm Dirac}
  =
  -\frac12\log\det P_{\mathcal D},
  \qquad
  \mathcal W_{\rm Maxwell}
  =
  \frac12\log\det P_1-\log\det P_0 .
\]
Here
\[
  P_{\mathcal D}=\mathcal D^2=-\nabla^2+\frac14R,
\]
while \(P_1\) is the Hodge Laplacian on one-forms and \(P_0\) the scalar
ghost Laplacian.  Gamma-matrix traces in the spin bundle and the
one-form/ghost bundle traces reduce the logarithmic terms to the following
table:
\[
\begin{array}{c|cc}
\text{free conformal field} & a & c\\ \hline
\text{real scalar} & \frac{1}{360} & \frac{1}{120}\\[0.2em]
\text{Weyl fermion} & \frac{11}{720} & \frac{1}{40}\\[0.2em]
\text{Dirac fermion} & \frac{11}{360} & \frac{1}{20}\\[0.2em]
\text{Maxwell vector} & \frac{31}{180} & \frac{1}{10}
\end{array}
\]
The Weyl fermion entries are half the Dirac entries.  The vector entries
include the Faddeev-Popov ghost determinant and refer to the gauge-invariant
stress tensor rather than the unconstrained one-form determinant alone.

\section{Example: N=4 Yang-Mills}

Consider free four-dimensional \(\mathcal N=4\) Yang-Mills fields in the
adjoint representation of a compact gauge group \(G\), with
\[
  \dim G=d_G.
\]
Per generator of \(G\), the field content is one Maxwell vector, four Weyl
fermions, and six real conformal scalars.  Therefore
\[
  a_{\mathcal N=4}/d_G
  =
  \frac{31}{180}
  +4\frac{11}{720}
  +6\frac1{360}
  =
  \frac14,
\]
and
\[
  c_{\mathcal N=4}/d_G
  =
  \frac1{10}
  +4\frac1{40}
  +6\frac1{120}
  =
  \frac14.
\]
Thus
\[
  a_{\mathcal N=4}=c_{\mathcal N=4}=\frac14\,\dim G
\]
in the free normalization.  Supersymmetry and current multiplet arguments
protect this equality in the interacting conformal theory once the theory
and stress-tensor multiplet have been constructed with a compatible
renormalization scheme.

\section{Contact Terms and Local Covariance}

The anomaly is a statement about coincident stress-tensor insertions as well
as one-point functions.  Functional differentiation of
\eqref{eq:curved-trace-anomaly-definition} gives
\[
  g_{\mu\nu}(x)
  \left\langle
  T^{\mu\nu}(x)T^{\rho\sigma}(y)
  \right\rangle_{\rm conn}
  =
  \frac{2}{\sqrt g(y)}
  \frac{\delta}{\delta g_{\rho\sigma}(y)}
  \mathcal A[g](x)
  +\text{index-variation contact terms}.
\]
At separated points in a flat conformal theory the trace vanishes.  The
curvature anomaly records the contact distribution obtained by varying
through curved backgrounds and then returning to flat space.  This is why
the same coefficient \(c\) controls both the \(W^2\) anomaly and the
normalization of the separated stress-tensor two-point function, whereas
\(a\) first appears in genuinely three-stress-tensor contact data.

In the locally covariant language, the assignment
\[
  (M,g,\lambda)\longmapsto \mathcal W[g,\lambda]
\]
should be replaced by a functorial algebra of observables, a class of
Hadamard states, and locally covariant composite fields.  The anomaly is then
the obstruction to choosing the stress tensor so that it is simultaneously
conserved, locally covariant, and Weyl covariant with the declared finite
renormalization freedom.

\section{Status of the Curved-Background Theorem}

For free scalar, spinor, and gauge fields, the heat-kernel and point-splitting
constructions give theorem-level anomaly formulas once the operator domain,
gauge fixing, ghost determinant, and finite local counterterms are fixed.  In
perturbative algebraic QFT, the anomaly appears in the local covariance and
scaling behavior of time-ordered products.  For a general nonperturbative
interacting QFT, the following ingredients are required:
\begin{enumerate}[label=(\roman*)]
\item a locally covariant curved-background QFT object replacing
  \(\mathcal W[g,\lambda]\);
\item stress-tensor insertions as locally covariant composite fields;
\item a contact-term prescription for multiple metric variations;
\item a proof that Weyl variation is local modulo allowed counterterms;
\item compatibility with gauge and gravitational anomaly lines for other
  background fields;
\item a rule identifying which finite curvature counterterms are part of the
  chosen renormalization scheme.
\end{enumerate}

\begin{openproblem}[Trace anomaly from nonperturbative QFT data]
\label{op:trace-anomaly-nonperturbative-data}
Formulate trace anomalies directly from nonperturbative local QFT data on
curved backgrounds.  The construction should identify the algebraic,
functorial, or analytic object replacing \(\mathcal W[g,\lambda]\), define
stress-tensor contact terms, prove locality of Weyl variation, and recover
the local Weyl cohomology class without importing a perturbative regulator as
part of the definition.
\end{openproblem}
