\chapter{The Hawking Effect}
\label{ch:curved-hawking-effect}

\ConventionsLorentzian

The Hawking effect is a statement about quantum fields on a fixed spacetime
whose causal structure contains a future black-hole horizon.  Its invariant
content is not that a preferred global particle basis exists.  Rather, under
regularity at the future horizon and suitable propagation to future null
infinity, late outgoing observables have a thermal occupation factor with
temperature fixed by the surface gravity.  This chapter separates four
ingredients which are often compressed into one slogan: horizon regularity,
exponential ray tracing, Bogoliubov mixing of wave packets, and propagation
through the exterior geometry.

\section{Stationary Horizons and Surface Gravity}

Let \((M,g)\) be a smooth Lorentzian spacetime and let \(\chi\) be a Killing
field.  A null hypersurface \(\mathcal H\) is a Killing horizon for \(\chi\)
if \(\chi\) is tangent to \(\mathcal H\), nonzero on the open horizon, and
null on \(\mathcal H\).  On \(\mathcal H\), the acceleration of the horizon
generators is necessarily proportional to \(\chi\):
\begin{equation}
  \chi^\nu\nabla_\nu\chi^\mu=\kappa\chi^\mu .
  \label{eq:hawking-surface-gravity-definition}
\end{equation}
The proportionality factor \(\kappa\) is the surface gravity in the
normalization of \(\chi\).  For an asymptotically flat stationary black hole
the normalization is fixed by requiring that \(\chi\) approach a unit time
translation at infinity.

\begin{lemma}[Rindler normal form of a nonextremal horizon]
\label{lem:hawking-rindler-normal-form}
Suppose \(\mathcal H\) is a nonextremal Killing horizon, \(\kappa>0\).  In a
neighborhood of a horizon cross-section, the two-dimensional metric in the
normal directions can be written, up to terms smooth in \(\rho^2\), as
\[
  ds^2_{\perp}
  =
  -\kappa^2\rho^2\,dt^2+d\rho^2 .
\]
Equivalently, the coordinates
\[
  U=-\rho e^{-\kappa t},\qquad V=\rho e^{\kappa t}
\]
are regular null coordinates through the future horizon \(U=0\), and
\(\chi=\partial_t\).
\end{lemma}

\begin{proof}
Choose a smooth cross-section \(\Sigma\subset\mathcal H\).  In a neighborhood
of \(\Sigma\) one may use Gaussian null coordinates
\[
  (v,r,y^A),
  \qquad
  \mathcal H=\{r=0\},
  \qquad
  \chi=\partial_v ,
\]
where \(y^A\) are coordinates on \(\Sigma\), \(v\) is the Killing parameter
along the generators, and \(r\) is an affine parameter on null geodesics
transverse to the horizon.  The metric has the form
\[
  g
  =
  2\,dv\,dr
  -2\kappa r\,dv^2
  +2r h_A(r,y)\,dv\,dy^A
  +\gamma_{AB}(r,y)\,dy^A dy^B .
\]
The coefficient of \(r\,dv^2\) is fixed by
\eqref{eq:hawking-surface-gravity-definition}: on \(r=0\),
\(\nabla_\chi\chi=\kappa\chi\), and nonextremality means \(\kappa>0\).
Thus the normal two-dimensional part, at fixed \(y\), is
\[
  ds_\perp^2=2\,dv\,dr-2\kappa r\,dv^2+O(r^2dv^2)+O(r\,dv\,dy).
\]
Put
\[
  r=\frac{\kappa}{2}\rho^2,
  \qquad
  t=v-\frac{1}{2\kappa}\log r
  =
  v-\frac1\kappa\log\rho+\text{constant}.
\]
Then
\[
  dv=dt+\frac{dr}{2\kappa r}
  =
  dt+\frac{d\rho}{\kappa\rho},
\]
and the leading normal metric becomes
\[
\begin{aligned}
  2\,dv\,dr-2\kappa r\,dv^2
  &=
  2\left(dt+\frac{d\rho}{\kappa\rho}\right)\kappa\rho\,d\rho
  -\kappa^2\rho^2
   \left(dt+\frac{d\rho}{\kappa\rho}\right)^2  \\
  &=
  -\kappa^2\rho^2dt^2+d\rho^2 .
\end{aligned}
\]
The omitted coefficients are smooth functions of \(r=\kappa\rho^2/2\) and of
\(y\), so the correction terms are smooth in \(\rho^2\) in the
Killing-adapted chart.  Finally define
\[
  U=-\rho e^{-\kappa t},
  \qquad
  V=\rho e^{\kappa t}.
\]
Then \(UV=-\rho^2\) and
\[
  -\kappa^2\rho^2dt^2+d\rho^2=-\,dU\,dV ,
\]
up to the conventional factor used in writing a symmetric null product.
Thus \(U,V\) extend smoothly through \(U=0\), the future horizon, and
\(\partial_t=\partial_v=\chi\).
\end{proof}

For Schwarzschild of mass \(M\) in units \(G=\hbar=c=k_B=1\),
\[
  ds^2= -\left(1-\frac{2M}{r}\right)dt^2
  +\left(1-\frac{2M}{r}\right)^{-1}dr^2+r^2d\Omega_2^2 .
\]
Writing \(r=2M+\rho^2/(8M)+O(\rho^4/M^3)\) gives
\[
  ds^2_{\perp}
  =
  -\left(\frac{\rho}{4M}\right)^2dt^2+d\rho^2+O(\rho^4),
\]
so
\[
  \kappa_{\rm Schw}=\frac1{4M},
  \qquad
  T_{\rm Schw}=\frac1{8\pi M}.
\]

\section{Euclidean Regularity}

Analytically continue the stationary time by \(t=-\ii\tau\).  The normal
metric becomes
\[
  ds_E^2\simeq \kappa^2\rho^2d\tau^2+d\rho^2 .
\]
This is the flat metric in polar coordinates if and only if the angular
coordinate \(\kappa\tau\) has period \(2\pi\).  The Euclidean regularity
argument therefore fixes
\[
  \beta_H=\frac{2\pi}{\kappa},
  \qquad
  T_H=\frac{\kappa}{2\pi}.
\]

\begin{remark}[What Euclidean regularity proves]
\label{rem:hawking-euclidean-regularity-scope}
Assume a Euclidean QFT has been constructed on the smooth Euclidean black-hole
section, the analytic continuation to stationary Lorentzian correlators
exists, and the Euclidean time circle is the continuation of the normalized
Killing time.  Then the analytically continued stationary correlation
functions satisfy the KMS periodicity at inverse temperature
\(\beta_H=2\pi/\kappa\).  This statement does not by itself prove existence
of an interacting Lorentzian Hadamard state on a collapse spacetime.
\end{remark}

Smoothness at \(\rho=0\) identifies \(\kappa\tau\) as an angle.  Hence the
Euclidean Schwinger functions are periodic under
\(\tau\mapsto\tau+2\pi/\kappa\), with the usual ordering around the Euclidean
circle.  Under the assumed Osterwalder-Schrader or analytic-continuation
reconstruction for the stationary problem, Euclidean periodicity becomes the
KMS boundary relation for Lorentzian Killing-time correlation functions.
The final sentence records an implication that is not part of the argument:
Euclidean smoothness constructs a stationary thermal object, whereas collapse
requires a state empty on \(\mathcal I^-\), regular at the future horizon,
and propagated through a time-dependent geometry.

\section{Exponential Ray Tracing in Collapse}

The collapse derivation begins with the geometry of late outgoing null rays.
Let \(v\) be advanced time on past null infinity \(\mathcal I^-\), and let
\(u\) be retarded time on future null infinity \(\mathcal I^+\).  Let \(v_0\)
label the last escaping ray: rays with \(v<v_0\) reach \(\mathcal I^+\),
while rays with \(v>v_0\) fall through the future horizon.

\paragraph{Universal late-time ray tracing.}
Assume the collapse spacetime settles at late retarded time to a stationary
nonextremal black-hole exterior with future-horizon Kruskal coordinate
\[
  U=-C_H e^{-\kappa u}
\]
near \(\mathcal I^+\), with \(C_H>0\), and assume the past-null coordinate
\(v\) is a smooth affine parameter through the last escaping ray with
\[
  U=A(v-v_0)+O((v-v_0)^2),
  \qquad A>0 .
\]
Then for \(v<v_0\),
\[
  v_0-v=C e^{-\kappa u}+O(e^{-2\kappa u}),
  \qquad C=\frac{C_H}{A}>0,
\]
or equivalently
\[
  u(v)=-\frac1\kappa\log\frac{v_0-v}{C}+O(v_0-v).
\]
For an escaping ray close to the horizon, \(U<0\).  Put \(s=v_0-v>0\).  The
smoothness assumption at the last escaping ray gives
\[
  U=-As+O(s^2),
\]
with \(A>0\).  The horizon coordinate relation gives the same \(U\) as
\[
  -C_H e^{-\kappa u}
  =
  -As+O(s^2).
\]
Multiplying by \(-1/A\) gives
\[
  s-\frac{C_H}{A}e^{-\kappa u}=O(s^2).
\]
For sufficiently late \(u\), \(s\) is small and the inverse-function theorem
applied to \(s\mapsto s+O(s^2)\) yields
\[
  s=\frac{C_H}{A}e^{-\kappa u}+O(e^{-2\kappa u}).
\]
Taking the logarithm of
\(s=C e^{-\kappa u}(1+O(e^{-\kappa u}))\), with \(C=C_H/A\), gives
\[
  u=-\frac1\kappa\log\frac{s}{C}+O(s),
\]
which is the stated inverse expansion.

The constants \(A,C_H,C\) are nonuniversal; the exponential rate \(\kappa\)
is fixed by the regularity of the horizon coordinate.

\section{Bogoliubov Coefficients from Mode Tracing}

We carry out the coefficient calculation for a massless scalar field in the
geometric-optics approximation.  Angular variables and the exterior
potential are suppressed in this section; they reappear as greybody factors
below.  Normalize right-moving modes on a null line by
\[
  f_{\omega'}^{\rm in}(v)
  =
  \frac{1}{\sqrt{4\pi\omega'}}e^{-\ii\omega' v},
  \qquad
  p_{\omega}^{\rm out}(u)
  =
  \frac{1}{\sqrt{4\pi\omega}}e^{-\ii\omega u},
  \qquad
  \omega,\omega'>0 .
\]
Tracing \(p_\omega^{\rm out}\) backwards to \(\mathcal I^-\) gives, near the
singular late-time part responsible for particle creation,
\begin{equation}
  p_\omega(v)
  =
  \frac{1}{\sqrt{4\pi\omega}}
  \left(\frac{v_0-v}{C}\right)^{\ii\omega/\kappa}
  \theta(v_0-v)
  +\text{terms smooth across }v=v_0 .
  \label{eq:hawking-traced-mode-singular-part}
\end{equation}
The smooth terms have rapidly decaying Fourier transforms at large
\(\omega'\) and do not determine the universal thermal ratio.

Use the Klein-Gordon product on a null line,
\[
  (f,g)_{\mathcal I^-}
  =
  -\ii\int_{-\infty}^{\infty}
  f(v)\overleftrightarrow{\partial_v}\overline{g(v)}\,dv .
\]
The expansion
\[
  p_\omega
  =
  \int_0^\infty d\omega'\,
  \left(
  \alpha_{\omega\omega'}f_{\omega'}^{\rm in}
  +
  \beta_{\omega\omega'}\overline{f_{\omega'}^{\rm in}}
  \right)
\]
has
\[
  \alpha_{\omega\omega'}=(f_{\omega'}^{\rm in},p_\omega)_{\mathcal I^-},
  \qquad
  \beta_{\omega\omega'}=-(\overline{f_{\omega'}^{\rm in}},p_\omega)_{\mathcal I^-}.
\]
Putting \(x=v_0-v\) and \(a=\omega/\kappa\), the needed oscillatory boundary
values are
\[
  \int_0^\infty x^{\ii a}e^{-\ii\omega' x}\,dx
  =
  e^{-\ii\pi/2}e^{\pi a/2}\Gamma(1+\ii a)(\omega')^{-1-\ii a},
\]
\[
  \int_0^\infty x^{\ii a}e^{+\ii\omega' x}\,dx
  =
  e^{+\ii\pi/2}e^{-\pi a/2}\Gamma(1+\ii a)(\omega')^{-1-\ii a}.
\]
Equivalently, since \(\Gamma(1+\ii a)=\ii a\,\Gamma(\ii a)\), the singular
coefficients can be written
\[
  \alpha_{\omega\omega'}
  =
  A_{\omega\omega'}\,e^{+\pi\omega/(2\kappa)}
  \Gamma(\ii\omega/\kappa)(\omega')^{-\ii\omega/\kappa},
\]
\[
  \beta_{\omega\omega'}
  =
  -A'_{\omega\omega'}\,e^{-\pi\omega/(2\kappa)}
  \Gamma(\ii\omega/\kappa)(\omega')^{-\ii\omega/\kappa},
\]
where \(A_{\omega\omega'}\) and \(A'_{\omega\omega'}\) differ only by phases
and have common modulus
\[
  |A_{\omega\omega'}|
  =
  \frac{1}{2\pi\kappa}\sqrt{\frac{\omega}{\omega'}} .
\]
Therefore
\[
  \frac{|\alpha_{\omega\omega'}|^2}{|\beta_{\omega\omega'}|^2}
  =
  e^{2\pi\omega/\kappa}.
\]
Using
\[
  |\Gamma(\ii a)|^2=\frac{\pi}{a\sinh(\pi a)}
\]
gives the explicit negative-frequency density
\begin{equation}
  |\beta_{\omega\omega'}|^2
  =
  \frac{1}{2\pi\kappa\,\omega'}
  \frac{1}{e^{2\pi\omega/\kappa}-1}.
  \label{eq:hawking-continuum-beta-density}
\end{equation}

Equation~\eqref{eq:hawking-continuum-beta-density} is not yet a finite
particle number.  The factor \(1/\omega'\) is the density remnant of an
idealized stationary late-time limit.  Finite observables are wave-packet
observables.

\section{Wave Packets and the Planck Factor}

Fix a frequency resolution \(\varepsilon>0\).  For integers \(j\ge0\) and
\(n\in\mathbb Z\), define outgoing packets
\begin{equation}
  p_{jn}^{\rm out}
  =
  \frac{1}{\sqrt\varepsilon}
  \int_{j\varepsilon}^{(j+1)\varepsilon}
  d\omega\,
  e^{2\pi\ii n\omega/\varepsilon}p_\omega^{\rm out}.
  \label{eq:hawking-wave-packet-definition}
\end{equation}
The integer \(j\) labels a frequency bin and \(n\) labels retarded-time
localization, with center approximately \(u_n=2\pi n/\varepsilon\).

\begin{controlledapproximation}[Late-time packet occupation]
\label{ca:hawking-wave-packet-occupation}
In the stationary late-time approximation determined by
the exponential ray-tracing relation above, the number of incoming-vacuum
quanta in the outgoing packet \((j,n)\) is
\begin{equation}
  \langle N_{jn}\rangle
  =
  \frac1{\varepsilon}
  \int_{j\varepsilon}^{(j+1)\varepsilon}
  \frac{d\omega}{e^{\beta_H\omega}-1}
  +o(1)
  \qquad(n\to+\infty),
  \label{eq:hawking-packet-bin-average}
\end{equation}
where \(\beta_H=2\pi/\kappa\).  If the bin is narrow compared with the scale
on which the Planck factor varies, this is
\[
  \langle N_{jn}\rangle
  =
  \frac{1}{e^{\beta_H\omega_j}-1}
  +O(\varepsilon\,\partial_\omega n_\omega),
  \qquad \omega_j=(j+\tfrac12)\varepsilon .
\]

The packet Bogoliubov coefficient is
\[
  \beta_{jn,\omega'}
  =
  \frac1{\sqrt\varepsilon}
  \int_{j\varepsilon}^{(j+1)\varepsilon}
  d\omega\,e^{2\pi\ii n\omega/\varepsilon}
  \beta_{\omega\omega'} .
\]
The packet number is
\[
  \langle N_{jn}\rangle
  =
  \int_0^\infty d\omega'\,|\beta_{jn,\omega'}|^2 .
\]
Inserting the singular coefficient gives a double integral over
\(\omega_1,\omega_2\) and the \(\omega'\)-factor
\[
  \int_0^\infty
  \frac{d\omega'}{\omega'}
  (\omega')^{\ii(\omega_2-\omega_1)/\kappa}.
\]
With a large finite affine-time regulator this is the Fourier transform of a
long interval in \(\log\omega'\).  In the infinite stationary limit it is the
distribution
\begin{equation}
  2\pi\kappa\,\delta(\omega_1-\omega_2).
  \label{eq:hawking-log-frequency-delta}
\end{equation}
The phase \(e^{2\pi\ii n(\omega_1-\omega_2)/\varepsilon}\) then drops out,
and the remaining diagonal coefficient is exactly the Planck factor in
\eqref{eq:hawking-continuum-beta-density}.  Nonstationary corrections from
the \(O(e^{-2\kappa u})\) part of the ray-tracing map and from the smooth
part of the traced mode give the \(o(1)\) term for packets moved to late
retarded time.
\end{controlledapproximation}

The exact bin average in \eqref{eq:hawking-packet-bin-average} can be written
without approximation:
\[
  \frac1{\varepsilon}
  \int_{\omega_-}^{\omega_+}
  \frac{d\omega}{e^{\beta_H\omega}-1}
  =
  \frac{1}{\beta_H\varepsilon}
  \log
  \frac{1-e^{-\beta_H\omega_+}}
       {1-e^{-\beta_H\omega_-}},
  \qquad
  \omega_\pm=j\varepsilon,(j+1)\varepsilon .
\]

\section{Two-Dimensional Flux and the Moving-Mirror Model}

For a two-dimensional conformal scalar, the same exponential map determines
the stress-tensor flux.  Let \(p(u)\) denote the incoming null coordinate of
the ray reaching \(\mathcal I^+\) at retarded time \(u\):
\[
  v=p(u)=v_0-Ce^{-\kappa u}.
\]
For a CFT of central charge \(c\) in the incoming vacuum, the chiral stress
tensor transforms under the reparametrization \(v=p(u)\) as
\begin{equation}
  \langle T_{uu}\rangle
  =
  -\frac{c}{24\pi}\{p,u\},
  \qquad
  \{p,u\}=\frac{p'''}{p'}-\frac32\left(\frac{p''}{p'}\right)^2 .
  \label{eq:hawking-schwarzian-flux}
\end{equation}
For the exponential map,
\[
  \{v_0-Ce^{-\kappa u},u\}=-\frac{\kappa^2}{2},
\]
and hence
\begin{equation}
  \langle T_{uu}\rangle
  =
  \frac{c\,\kappa^2}{48\pi}
  =
  \frac{\pi c}{12\beta_H^2}.
  \label{eq:hawking-two-dimensional-flux}
\end{equation}
The last equality is the thermal energy flux of one chiral CFT sector:
\[
  \int_0^\infty\frac{d\omega}{2\pi}
  \frac{\omega}{e^{\beta_H\omega}-1}
  =
  \frac{\pi}{12\beta_H^2}
\]
for \(c=1\).  The moving-mirror model realizes this calculation in flat
space by choosing a mirror trajectory whose ray-tracing function \(p(u)\)
has the same late exponential form.  It is a useful model of the kinematics
of Hawking radiation, not a derivation of quantum gravity.

\section{Trans-Planckian Precursor Frequencies}

The same tracing relation also shows the trans-Planckian domain of the
geometric-optics derivation.  A packet observed at retarded time \(u\) with
frequency \(\omega\) has precursor frequency on \(\mathcal I^-\) of order
\[
  \omega_{\rm in}
  \sim
  \omega\,\frac{du}{dv}
  =
  \frac{\omega}{\kappa(v_0-v)}
  \sim
  \frac{\omega}{\kappa C}e^{\kappa u}.
\]
For any fixed microscopic scale \(\Lambda_{\rm micro}\), sufficiently late
packets have \(\omega_{\rm in}>\Lambda_{\rm micro}\).  Thus the
fixed-background continuum derivation is a conditional theorem inside the
assumed QFT on the assumed smooth geometry.  Robustness under modified
short-distance physics is a separate statement requiring a specified
microscopic model and a controlled limit.

\section{Boulware, Hartle--Hawking, and Unruh States}

In a stationary eternal black-hole exterior, three quasifree scalar states
serve as reference points.
\begin{enumerate}[label=(\roman*)]
\item The Boulware state is positive-frequency with respect to the static
  Killing time at infinity.  It is empty for static observers at infinity but
  is singular at the horizon.
\item The Hartle--Hawking state is regular on both future and past horizons
  of the bifurcate extension and is KMS at inverse temperature
  \(\beta_H\) with respect to the exterior Killing time.
\item The Unruh state is regular on the future horizon and empty on past null
  infinity.  It is the fixed-background state appropriate to collapse and
  gives outgoing flux on \(\mathcal I^+\).
\end{enumerate}
These names denote state-selection properties, not extra axioms.  In a given
model, each state must be constructed as a positive Hadamard two-point
function or by another equivalent locally covariant construction.

\section{Greybody Factors and Four-Dimensional Flux}

In an asymptotically flat black-hole spacetime, propagation from the
near-horizon region to infinity includes scattering by the exterior curvature
potential.  For a scalar field on Schwarzschild, write the mode as
\[
  \phi_{\omega\ell m}
  =
  \frac{1}{r}\psi_{\omega\ell}(r_\ast)Y_{\ell m}(\theta,\varphi)
  e^{-\ii\omega t},
  \qquad
  \frac{dr_\ast}{dr}=\left(1-\frac{2M}{r}\right)^{-1}.
\]
The radial equation is
\[
  \left[
  -\frac{d^2}{dr_\ast^2}
  +V_\ell(r)
  \right]\psi_{\omega\ell}
  =
  \omega^2\psi_{\omega\ell},
  \qquad
  V_\ell(r)=
  \left(1-\frac{2M}{r}\right)
  \left(\frac{\ell(\ell+1)}{r^2}+\frac{2M}{r^3}\right).
\]
Let \(\Gamma_\ell(\omega)\) be the transmission probability through this
one-dimensional scattering problem, normalized by unit incoming flux from
the horizon.  The number and energy fluxes at infinity are
\[
  \frac{dN}{dt\,d\omega}
  =
  \sum_{\ell,m}
  \frac{\Gamma_\ell(\omega)}
  {2\pi(e^{\beta_H\omega}-1)},
\]
\[
  \frac{dE}{dt\,d\omega}
  =
  \sum_{\ell,m}
  \frac{\omega\,\Gamma_\ell(\omega)}
  {2\pi(e^{\beta_H\omega}-1)} .
\]
The Planck factor is horizon kinematics; \(\Gamma_\ell(\omega)\) is ordinary
hyperbolic propagation through the exterior.  Fermions replace the
Bose-Einstein denominator by \(e^{\beta_H\omega}+1\), and gauge fields
require the corresponding physical-polarization and ghost-free Hilbert-space
construction.

\begin{controlledapproximation}[Interacting horizon flux package]
\label{ca:hawking-interacting-horizon-flux-package}
For an interacting locally covariant QFT, the horizon part of the Hawking
argument should be phrased as a KMS spectral-density statement for renormalized
observables, not as a free particle-number formula.  Assume a stationary
late-time exterior, an admissible interacting Hadamard state \(\omega_U\)
regular at the future horizon, and a near-horizon algebra whose outgoing
channel \(A\) has bosonic Wightman boundary values
\[
  G_A^>(\omega),\qquad G_A^<(\omega),
  \qquad \omega>0 .
\]
The KMS condition at \(\beta_H=2\pi/\kappa\) gives
\[
  G_A^<(\omega)=e^{-\beta_H\omega}G_A^>(\omega),
  \qquad
  \rho_A(\omega):=G_A^>(\omega)-G_A^<(\omega)\ge0.
\]
Therefore
\begin{equation}
  G_A^<(\omega)=
  \frac{\rho_A(\omega)}{e^{\beta_H\omega}-1},
  \qquad
  G_A^>(\omega)=
  \frac{e^{\beta_H\omega}\rho_A(\omega)}
       {e^{\beta_H\omega}-1}.
  \label{eq:hawking-interacting-kms-spectral-density}
\end{equation}
The universal factor is the KMS ratio.  The interacting spectral density
\(\rho_A\), the channel mixing, the stress-tensor renormalization, and the
exterior propagation are model-dependent data.

On a retained frequency window \(I\) at future null infinity, a controlled
energy-flux statement has the form
\begin{equation}
  F_I^{\rm ret}
  =
  \sum_A
  \int_I\frac{d\omega}{2\pi}\,
  \omega\,\Gamma_A(\omega)\,
  \frac{\rho_A(\omega)}{e^{\beta_H\omega}-1}
  +
  R_I .
  \label{eq:hawking-interacting-retained-flux}
\end{equation}
Here \(\Gamma_A(\omega)\) is the exterior transmission probability or, more
generally, the positive propagation weight from the horizon channel to the
chosen null-infinity observable.  The residual \(R_I\) must include at least
the state-construction error, the KMS/late-time error, spectral-density and
operator-mixing truncation, greybody propagation error, stress-tensor
renormalization and Ward-identity residuals, nonstationary collapse tails, and
backreaction corrections.  A free scalar one-particle channel is recovered only
after choosing \(\rho_A=1\) in the corresponding normalization and taking the
other residuals to the declared limit.  In an interacting theory, replacing
\(\rho_A\) by the free value or fitting only the Bose denominator loses the
physical information carried by the local QFT.

This is also the correct interface with the semiclassical equation.  The mass
loss law uses the renormalized stress flux
\(F_I=\int_I\langle T_{uu}^{\rm ren}\rangle\), not a basis-dependent particle
count.  Particle language is justified only when the chosen channel basis,
state, propagation map, and stress-tensor conversion have all been controlled
on the same window.
\end{controlledapproximation}

\section{Back-Reaction Boundary}

Fixed-background QFT treats \(g\) as prescribed.  Semiclassical gravity asks
instead for
\[
  G_{\mu\nu}[g]+\Lambda g_{\mu\nu}
  =
  8\pi G\,\langle T_{\mu\nu}\rangle_{\rm ren}[g,\omega],
\]
where \(\omega\) is the chosen Hadamard state and the stress tensor includes
the finite local curvature counterterms discussed earlier in this volume.  In
an adiabatic Schwarzschild approximation, energy conservation suggests an
ODE
\[
  \frac{dM}{du}=-L(M),
  \qquad
  L(M)=
  \sum_{\ell,m}
  \int_0^\infty
  \frac{d\omega}{2\pi}\,
  \frac{\omega\,\Gamma_\ell(\omega;M)}
       {e^{8\pi M\omega}-1},
\]
for a scalar contribution.  This equation is not a theorem about the full
evaporating geometry.  It assumes slow variation on the scale \(M\), ignores
stress-tensor fluctuations except through their mean, and presupposes that
the state remains close to an Unruh-type Hadamard state on each instantaneous
exterior.

\begin{controlledapproximation}[Flux-to-mass backreaction window]
\label{ca:hawking-flux-to-mass-window}
Fix a retarded-time interval \(J=[u_0,u_0+\Delta u]\) during which the
geometry is represented by a Schwarzschild parameter \(M(u)\) in a chart
\(|M(u)-M_0|\leq \delta M\).  Let \(L_I^{\rm ret}(M_0)\) be the retained
renormalized stress-flux luminosity obtained from
\eqref{eq:hawking-interacting-retained-flux}, including the selected
frequency window and channels.  The mean mass-loss approximation is the
finite-window statement
\[
  M(u_0+\Delta u)
  =
  M_0-\Delta u\,L_I^{\rm ret}(M_0)
  +R_{\rm flux}+R_{\rm drift}+R_{\rm state}+R_{\rm grav}.
\]
These are signed remainders whose absolute values must be bounded by declared
nonnegative error budgets.  The residuals have distinct meanings.
\(R_{\rm flux}\) is the spectral, greybody, stress-renormalization, Ward, and
tail budget already attached to the retained flux.  \(R_{\rm drift}\) is the
error from replacing \(L_I(M(u))\) by \(L_I(M_0)\).  If \(L_I\) is Lipschitz
with constant \(K_I\) on the chart, then
\[
  |R_{\rm drift}|
  \leq
  \Delta u\,K_I\,\delta M .
\]
\(R_{\rm state}\) measures failure of the transported state to remain in the
declared Unruh-type Hadamard class on the window, and \(R_{\rm grav}\)
measures omitted gravitational EFT, higher-derivative, and nonlinear metric
corrections.

The same window must also control fluctuations.  Let
\[
  \mathcal N_J
  =
  \int_J du\int_J du'\,
  \langle \delta F(u)\delta F(u')\rangle_{\rm c}
\]
be the retained connected covariance of the renormalized flux observable
\(F(u)=\int_{\mathcal I^+_u}T_{uu}^{\rm ren}\), with its own nonnegative
residual \(R_{\mathcal N}\).  A mean evaporating geometry is a controlled
output only when
\[
  \epsilon_{\rm qs}
  :=
  \Delta u\,
  \frac{L_I^{\rm ret}(M_0)+|R_{\rm flux}|/\Delta u}
       {M_0}
  \ll 1,
  \qquad
  \sqrt{\mathcal N_J+R_{\mathcal N}}\ll \delta M .
\]
The first inequality is the quasi-stationary condition
\(\Delta u\,|\dot\kappa|/\kappa\ll1\) for \(\kappa=1/(4M)\).  The second says
that the integrated flux noise remains inside the same metric chart in which
the mean mass loss is being computed.  If either test fails, the stationary
Hawking flux remains a fixed-background observable, but the adiabatic
evaporating metric has not been derived.
\end{controlledapproximation}

\section{QFT Boundary}

A theorem-level Hawking effect for an interacting local QFT on black-hole
backgrounds requires all of the following data:
\begin{enumerate}[label=(\roman*)]
\item a precise class of globally hyperbolic collapse spacetimes with
  controlled late approach to a nonextremal stationary exterior;
\item construction of a Hadamard state empty on \(\mathcal I^-\) and regular
  at the future horizon;
\item a local and covariant stress-tensor renormalization scheme;
\item propagation estimates connecting near-horizon data to \(\mathcal I^+\);
\item control of the limits defining late-time wave-packet observables;
\item a statement separating the universal horizon factor from
  model-dependent greybody and interaction effects.
\end{enumerate}
Without these hypotheses, the phrase ``black holes radiate thermally'' is not
a mathematical theorem.  With these hypotheses supplied, the fixed-background
calculation above shows where the temperature and Planck factor enter.

\begin{openproblem}[Interacting Hawking theorem]
\label{op:interacting-hawking-theorem}
Prove a Hawking radiation theorem for an interacting local QFT on a class of
black-hole backgrounds, including Hadamard state construction, KMS or
near-horizon analyticity, renormalized stress tensor, greybody propagation,
late-time wave-packet limits, and the precise sense in which the emitted
flux is thermal.
\end{openproblem}
