\chapter{Point Splitting and Stress Tensor Renormalization}
\label{ch:point-splitting-stress-tensor-renormalization}

\ConventionsMixed

Chapter~\ref{ch:locally-covariant-qft-hadamard} introduced Hadamard states as
the admissible short-distance class for free fields on curved backgrounds,
after the microlocal wavefront-set criterion was fixed in
Chapter~\ref{ch:curved-microlocal-spectrum-condition}.  Point splitting uses
the state-independent singular part of a Hadamard two-point function to define
composite local fields.
The exact upstream inputs are the microlocal Hadamard two-point condition,
the local Hadamard parametrix and coefficient recursion, and the product
criterion for distributional contractions away from forbidden wavefront
directions.  What remains in this chapter is the finite local choice of a
composite field, especially the stress tensor.

\section{Local Hadamard Subtraction Datum}
\label{sec:point-splitting-local-hadamard-subtraction-datum}

Throughout this chapter the scalar operator is the one fixed in
Chapter~\ref{ch:locally-covariant-qft-hadamard},
\begin{equation}
  P_M=-\nabla^\mu\nabla_\mu+m^2+\xi R[g],
  \qquad m^2\ge0,\qquad \xi\in\mathbb R .
  \label{eq:point-splitting-kg-operator}
\end{equation}
Let \(\mathcal U\subset M\) be a geodesically convex normal neighborhood.
For \(x,y\in\mathcal U\), let \(\sigma(x,y)\) be Synge's world function:
one half of the signed squared geodesic distance from \(x\) to \(y\).  Choose
a smooth time function \(T\) on \(\mathcal U\) and put
\begin{equation}
  \sigma_\epsilon(x,y)
  =
  \sigma(x,y)+\ii\epsilon\bigl(T(x)-T(y)\bigr)+\epsilon^2,
  \qquad \epsilon>0 .
  \label{eq:point-splitting-sigma-epsilon}
\end{equation}
The limit \(\epsilon\downarrow0\) is always distributional.

\begin{definition}[Hadamard subtraction]
\label{def:local-hadamard-subtraction}
In four dimensions, a Hadamard subtraction for \(P_M\) on
\(\mathcal U\) is the distributional boundary value
\begin{equation}
  H_{\epsilon,\mu}(x,y)
  =
  \frac{1}{8\pi^2}
  \left[
    \frac{U_H(x,y)}{\sigma_\epsilon(x,y)}
    +V_H(x,y)\log\!\bigl(\mu^2\sigma_\epsilon(x,y)\bigr)
  \right],
  \qquad \mu>0,
  \label{eq:point-splitting-hadamard-subtraction}
\end{equation}
where \(U_H=\Delta^{1/2}\), \(\Delta\) is the van Vleck determinant, and
\(V_H\) is a smooth representative of the Hadamard coefficient whose jet at
the diagonal is determined recursively by \(P_M\).  Different smooth
representatives with the same diagonal jet change no finite point-split
quantity of bounded differential order.  Additional smooth local biscalars
may be added only as locally covariant finite renormalizations; they are not
part of the state.
\end{definition}

The word ``determined'' in
Definition~\ref{def:local-hadamard-subtraction} means a local transport
construction of the jet of \(V_H\) at the diagonal.  It is not a global
Green-function prescription.  To spell out the local object, write
\begin{equation}
  V_H(x,y)\sim \sum_{j\ge0}v_j(x,y)\sigma(x,y)^j
  \label{eq:hadamard-v-asymptotic-expansion}
\end{equation}
as an asymptotic expansion of smooth biscalars in a normal neighborhood.
For each integer \(N\ge0\), the partial sum
\begin{equation}
  V_H^{[N]}(x,y)=\sum_{j=0}^{N}v_j(x,y)\sigma(x,y)^j
  \label{eq:hadamard-v-partial-sum}
\end{equation}
is chosen so that
\begin{equation}
  P_{M,x}\left[
    \frac{U_H(x,y)}{\sigma(x,y)}
    +
    V_H^{[N]}(x,y)\log\!\bigl(\mu^2\sigma(x,y)\bigr)
  \right]
  \label{eq:hadamard-recursion-condition}
\end{equation}
has no negative powers of \(\sigma\) and no logarithmic terms through the
order fixed by \(N\).  Expanding
\eqref{eq:hadamard-recursion-condition} in powers of \(\sigma\) gives
ordinary differential equations along the unique geodesic from \(y\) to
\(x\).  Since the coefficient multiplying the unknown diagonal value at the
\(j\)-th step is nonzero, the transport equation fixes \(v_j\) from lower
coefficients.  A smooth \(V_H\) with this prescribed infinite jet exists by
Borel's lemma.  A natural locally covariant subtraction requires, in addition,
a rule compatible with metric embeddings and parameter dependence.  Replacing
one extension by another one with the same jet changes \(H_{\epsilon,\mu}\) by
a smooth biscalar flat at the diagonal before multiplication by
\(\log\sigma\), so any point-split composite field using finitely many
diagonal derivatives is unchanged.  A nonflat smooth addition is a genuine
renormalization choice and must be built locally and covariantly from
\(g_{\mu\nu}\), \(m^2\), \(\xi\), and their curvature data.

\begin{lemma}[Leading Hadamard transport equation]
\label{lem:leading-hadamard-transport-equation}
The coefficient \(U_H\) in
\eqref{eq:point-splitting-hadamard-subtraction} is fixed by
\begin{equation}
  2\sigma^{;\mu}\nabla_\mu U_H
  +
  (\Box_g\sigma-4)U_H=0,
  \qquad U_H(x,x)=1 .
  \label{eq:point-splitting-u-transport}
\end{equation}
\end{lemma}

\begin{proof}
The calculation is local and may be performed for \(x\ne y\), before taking
the distributional boundary value.  The defining identities of Synge's
function in a normal neighborhood are
\begin{equation}
  \sigma^{;\mu}\sigma_{;\mu}=2\sigma,\qquad
  \Box_g\sigma=4+O(\sigma).
  \label{eq:point-splitting-synge-identities}
\end{equation}
For a smooth biscalar \(U\),
\[
  \nabla_\mu(\sigma^{-1})=-\sigma_{;\mu}\sigma^{-2},
  \qquad
  \Box_g(\sigma^{-1})
  =
  -(\Box_g\sigma)\sigma^{-2}
  +2\sigma^{;\mu}\sigma_{;\mu}\sigma^{-3}.
\]
Using \(\sigma^{;\mu}\sigma_{;\mu}=2\sigma\), this becomes
\[
  \Box_g(\sigma^{-1})=(4-\Box_g\sigma)\sigma^{-2}.
  \]
Hence the most singular term in
\(-\Box_{g,x}(U/\sigma)\) is
\[
  \frac{
    2\sigma^{;\mu}\nabla_\mu U
    +(\Box_g\sigma-4)U
  }{\sigma^2}.
\]
The zeroth-order part \(m^2+\xi R\) of \(P_M\) contributes only
\(\sigma^{-1}\) singularities, so the \(\sigma^{-2}\) coefficient must
vanish.  This is exactly \eqref{eq:point-splitting-u-transport}.  The
initial value \(U_H(x,x)=1\) gives the flat-space normalization and uniquely
solves the transport equation along the geodesic from \(y\) to \(x\); the
solution is \(\Delta^{1/2}\).
\end{proof}

\begin{lemma}[First logarithmic Hadamard coefficient]
\label{lem:first-logarithmic-hadamard-coefficient}
Let \(V_H=v_0+O(\sigma)\).  With the operator convention
\eqref{eq:point-splitting-kg-operator}, the first coefficient is fixed by
\begin{equation}
  2\sigma^{;\mu}\nabla_\mu v_0
  +
  (\Box_g\sigma-2)v_0
  =
  P_{M,x}U_H .
  \label{eq:point-splitting-v0-transport}
\end{equation}
In particular,
\begin{equation}
  v_0(x,x)
  =
  \frac12\left(m^2+\left(\xi-\frac16\right)R[g](x)\right).
  \label{eq:point-splitting-v0-coincidence}
\end{equation}
\end{lemma}

\begin{proof}
After Lemma~\ref{lem:leading-hadamard-transport-equation}, the
\(\sigma^{-2}\) coefficient in \(P_{M,x}(U_H/\sigma)\) vanishes.  The next
singular term is therefore
\begin{equation}
  \frac{P_{M,x}U_H}{\sigma}.
  \label{eq:point-splitting-pu-over-sigma}
\end{equation}
For a smooth biscalar \(v\),
\begin{equation}
  \nabla_\mu\log\sigma=\frac{\sigma_{;\mu}}{\sigma},
  \qquad
  \Box_g\log\sigma
  =
  \frac{\Box_g\sigma-2}{\sigma},
  \label{eq:point-splitting-box-log-sigma}
\end{equation}
where the second identity uses
\(\sigma^{;\mu}\sigma_{;\mu}=2\sigma\).  Hence
\begin{equation}
  P_{M,x}\bigl(v\log\sigma\bigr)
  =
  (P_{M,x}v)\log\sigma
  -
  \frac{
    2\sigma^{;\mu}\nabla_\mu v
    +(\Box_g\sigma-2)v
  }{\sigma}.
  \label{eq:point-splitting-p-vlog}
\end{equation}
The zeroth-order part of \(P_M\) multiplies \(v\log\sigma\) and contributes
only the displayed logarithmic term, not a negative power of \(\sigma\).
Canceling the coefficient of \(\sigma^{-1}\) between
\eqref{eq:point-splitting-pu-over-sigma} and
\eqref{eq:point-splitting-p-vlog} gives
\eqref{eq:point-splitting-v0-transport}.  Taking the coincidence limit,
\(\sigma^{;\mu}=0\) and \(\Box_g\sigma=4\).  Also
\begin{equation}
  U_H(x,x)=1,\qquad
  \left.\Box_{g,x}U_H(x,y)\right|_{y=x}=\frac16R[g](x),
  \label{eq:point-splitting-delta-box-coincidence}
\end{equation}
which follows from the normal-coordinate expansion of
\(\Delta^{1/2}\).  Therefore
\[
  \left.P_{M,x}U_H(x,y)\right|_{y=x}
  =
  m^2+\xi R-\frac16R,
\]
and \eqref{eq:point-splitting-v0-coincidence} follows.
\end{proof}

\begin{proposition}[Hadamard logarithmic recursion]
\label{prop:hadamard-logarithmic-recursion}
Let
\[
  V_H(x,y)\sim\sum_{j\ge0}v_j(x,y)\sigma(x,y)^j
\]
be the local Hadamard logarithmic coefficient in four dimensions.  After
\(U_H\) has been fixed by
\eqref{eq:point-splitting-u-transport}, the coefficients \(v_j\) are
determined recursively along the geodesic from \(y\) to \(x\) by
\begin{equation}
  \bigl(2\sigma^{;\mu}\nabla_\mu+\Box_g\sigma-2\bigr)v_0
  =
  P_{M,x}U_H,
  \label{eq:hadamard-recursion-v0}
\end{equation}
and, for \(j\ge0\),
\begin{equation}
  \bigl(2\sigma^{;\mu}\nabla_\mu+\Box_g\sigma+2j\bigr)v_{j+1}
  =
  \frac{1}{j+1}\,P_{M,x}v_j .
  \label{eq:hadamard-recursion-vj}
\end{equation}
At coincidence this gives
\begin{equation}
  v_{j+1}(x,x)
  =
  \frac{1}{2(j+1)(j+2)}
  \left.P_{M,x}v_j(x,y)\right|_{y=x}.
  \label{eq:hadamard-recursion-coincidence}
\end{equation}
The recursion fixes the complete diagonal jet of \(V_H\).  Choosing a smooth
representative with that jet is a Borel-extension choice and changes the
parametrix only by a smooth biscalar flat at the diagonal, hence only by a
finite local subtraction in point-split composite fields.
\end{proposition}

\begin{proof}
Write \(s=\sigma(x,y)\) and compute away from the diagonal.  For a smooth
biscalar \(v\),
\begin{equation}
  P_{M,x}(v\,s^k\log s)
  =
  \bigl(P_{M,x}(v\,s^k)\bigr)\log s
  -
  s^{k-1}
  \bigl(
    2\sigma^{;\mu}\nabla_\mu v
    +(\Box_g\sigma+4k-2)v
  \bigr).
  \label{eq:hadamard-recursion-basic-identity}
\end{equation}
This follows by applying the product rule and
\(\sigma^{;\mu}\sigma_{;\mu}=2\sigma\), exactly as in the proof of
Lemma~\ref{lem:first-logarithmic-hadamard-coefficient}.  The
zeroth-order part \(m^2+\xi R\) contributes to the logarithmic term
\((P_{M,x}(v\,s^k))\log s\), but not to the displayed nonlogarithmic
coefficient.

The \(\sigma^{-1}\) nonlogarithmic singularity in
\[
  P_{M,x}\left(\frac{U_H}{s}+v_0\log s\right)
\]
is \(P_{M,x}U_H\) minus the \(k=0\) coefficient in
\eqref{eq:hadamard-recursion-basic-identity}; setting it to zero gives
\eqref{eq:hadamard-recursion-v0}.  For \(j\ge0\), inspect the coefficient of
\(s^j\log s\) in
\[
  P_{M,x}\left(\sum_{\ell\ge0}v_\ell s^\ell\right).
\]
The term \(P_{M,x}(v_j s^j)\) contributes \(s^jP_{M,x}v_j\), while the
lower-derivative part of \(P_{M,x}(v_{j+1}s^{j+1})\) contributes
\[
  -(j+1)s^j
  \bigl(
    2\sigma^{;\mu}\nabla_\mu v_{j+1}
    +(\Box_g\sigma+2j)v_{j+1}
  \bigr).
\]
Vanishing of the coefficient of \(s^j\log s\) gives
\eqref{eq:hadamard-recursion-vj}.  At coincidence
\(\sigma^{;\mu}=0\) and \(\Box_g\sigma=4\), so
\eqref{eq:hadamard-recursion-vj} gives
\eqref{eq:hadamard-recursion-coincidence}.  The transport operator has
nonzero coincidence coefficient at each step, so the value at \(x=y\) and
then all transverse derivatives are fixed recursively by differentiating the
transport equation.  Borel's lemma supplies a smooth biscalar with the
prescribed infinite jet, and two such choices differ by a smooth biscalar
flat at the diagonal.
\end{proof}

Changing the time function in \eqref{eq:point-splitting-sigma-epsilon}
changes only the boundary-value prescription for the same local positive
frequency singularity.  Changing \(\mu\) changes the subtraction by the
smooth local biscalar
\begin{equation}
  \frac{1}{8\pi^2}V_H(x,y)\log(\mu'^2/\mu^2).
  \label{eq:point-splitting-scale-change}
\end{equation}
Thus the singular part is state-independent, while the scale and smooth
local terms are renormalization coordinates.

\section{Wick Square}

\begin{definition}[Wick square by point splitting]
\label{def:wick-square-point-splitting}
Let \(\omega\) be a Hadamard state of the real Klein--Gordon field.  Let
\(H(x,y)\) denote the locally covariant Hadamard parametrix, containing the
singular \(U/\sigma_\epsilon\) and \(V\log\sigma_\epsilon\) terms.  The
point-split Wick square in state \(\omega\) is the distribution defined by
\[
  \omega\bigl(:\!\Phi^2\!:(f)\bigr)
  =
  \int_M f(x)
  \lim_{y\to x}
  \bigl(\omega_2(x,y)-H(x,y)\bigr)\,d{\rm vol}_g(x).
\]
\end{definition}

\begin{remark}[State dependence and subtraction dependence]
\label{rem:point-splitting-smooth-change}
Fix a normal convex neighborhood in which the Hadamard parametrix \(H\) is
defined.  If \(\omega\) and \(\omega'\) are Hadamard states of the same
Klein--Gordon field, then
\[
  W_{\omega,\omega'}(x,y)
  =
  \omega_2(x,y)-\omega'_2(x,y)
\]
is smooth near the diagonal, and the corresponding Wick-square expectation
values obey
\[
  \omega\bigl(:\!\Phi^2\!:_H(f)\bigr)
  -
  \omega'\bigl(:\!\Phi^2\!:_H(f)\bigr)
  =
  \int_M f(x)\,W_{\omega,\omega'}(x,x)\,d{\rm vol}_g(x).
\]
If \(H'=H+S\), where \(S(x,y)\) is a smooth symmetric locally constructed
biscalar, then
\[
  :\!\Phi^2\!:_{H'}(f)
  =
  :\!\Phi^2\!:_{H}(f)
  -
  \int_M f(x)S(x,x)\,d{\rm vol}_g(x)\,\mathbf 1
\]
as a relation between expectation values in every Hadamard state.
\end{remark}

The defining property of a Hadamard two-point function is that, in each
normal convex neighborhood, its singular part is the universal local
parametrix \(H\), while the remainder is smooth.  Thus
\[
  \omega_2=H+R_\omega,\qquad
  \omega'_2=H+R_{\omega'}
\]
with \(R_\omega\) and \(R_{\omega'}\) smooth near the diagonal.  Their
difference \(W_{\omega,\omega'}=R_\omega-R_{\omega'}\) is smooth, so the
diagonal value \(W_{\omega,\omega'}(x,x)\) is an ordinary smooth function.
Subtracting the two point-splitting definitions gives the first displayed
formula.  For the second formula,
\[
  \omega_2-H'=\omega_2-H-S .
\]
Taking the diagonal limit and smearing with \(f\) subtracts
\(\int f(x)S(x,x)d{\rm vol}_g(x)\), independent of the state.  A
state-independent scalar multiple of every expectation value is represented
by the same scalar multiple of the identity field.

For higher Wick powers, one subtracts the parametrix in every contraction.
The result is an operator-valued distribution in perturbative algebraic QFT
and a quadratic form in representations where the required domains are
available.

The subtraction has finite local freedom.  For the Wick square in four
dimensions,
\[
  :\!\Phi^2\!:\ \mapsto\ :\!\Phi^2\!:+c_1R+c_2m^2
\]
with constants \(c_1,c_2\), subject to the chosen covariance and scaling
conditions.

\section{Stress-Tensor Prescription and Local Freedom}
\label{sec:stress-tensor-prescription-local-freedom}

The stress tensor has a sharper renormalization problem than the Wick square
because a choice of finite local terms must also respect conservation and
background covariance.

\begin{definition}[Wald-type stress-tensor renormalization conditions]
\label{def:wald-type-stress-tensor-conditions}
For the scalar theory with operator \(P_M\) in
\eqref{eq:point-splitting-kg-operator}, an expectation-value stress-tensor
renormalization is an assignment
\[
  (M,g;m^2,\xi,\omega)\longmapsto
  \omega(T_{\mu\nu}^{\rm ren})
  \in\mathcal D'(M,S^2T^*M)
\]
for every globally hyperbolic spacetime \((M,g)\) and every Hadamard state
\(\omega\).  In this chapter the renormalization conditions are the
following.

\begin{enumerate}
\item \emph{Local covariance.}  If
\(\chi:(M,g)\to(N,h)\) is an orientation- and time-orientation-preserving
isometric embedding with causally convex image, and if \(\omega_N\) is a
Hadamard state on \(N\), then the expectation value in the pulled-back state
\(\chi^*\omega_N\) equals the pullback of the expectation value on \(N\).
\item \emph{State dependence only through the smooth remainder.}  For any
two Hadamard states \(\omega,\omega'\), the difference
\(\omega(T_{\mu\nu}^{\rm ren})-\omega'(T_{\mu\nu}^{\rm ren})\) is obtained
by applying a fixed local bidifferential operator to the smooth biscalar
\(\omega_2-\omega'_2\) and then restricting to the diagonal.
\item \emph{Conservation.}  The distributional divergence vanishes,
\begin{equation}
  \nabla^\mu\omega(T_{\mu\nu}^{\rm ren})=0 ,
  \label{eq:wald-type-conservation}
\end{equation}
for all Hadamard states in the declared state class.
\item \emph{Locality, weak regularity, and almost homogeneous scaling.}  In
local formulae we assign engineering dimensions
\begin{equation}
  [\nabla]=1,\qquad [m]=1,\qquad [R_{\mu\nu\rho\sigma}]=2,\qquad
  [g_{\mu\nu}]=0 .
  \label{eq:stress-tensor-engineering-dimensions}
\end{equation}
The state-independent local terms in \(T_{\mu\nu}^{\rm ren}\) are natural
tensors of total dimension four, depend smoothly on compactly supported
metric perturbations and on the parameters \(m^2,\xi\), and scale almost
homogeneously under constant rescalings of the metric, with only a finite
logarithmic scaling defect.  These regularity hypotheses are part of the
condition: dimension counting alone would allow, for fixed \(m>0\), tensors
from nonpolynomial functionals such as
\(\int\sqrt{|g|}\,m^4F(R/m^2)\).  Smooth parameter dependence at \(m^2=0\),
local covariance, and almost homogeneous scaling reduce the allowed
dimension-four local tensors to the finite list classified below.  A change
of the Hadamard length scale is allowed to produce only local logarithmic
terms of the same dimension; this is the curved-space form of the
trace-anomaly ambiguity.
\item \emph{Flat normalization.}  On Minkowski spacetime the massless
vacuum expectation value is zero after the finite cosmological term is set
to zero.
\end{enumerate}
These are renormalization conditions on expectation values.  A stronger
operator-valued statement additionally requires a common domain or an
algebraic definition of the composite field; that extra domain problem is
not hidden in the word ``stress tensor''.
\end{definition}

\begin{definition}[Point-split stress-tensor prescription]
\label{def:point-split-stress-tensor-prescription}
A point-split realization of
Definition~\ref{def:wald-type-stress-tensor-conditions} for the scalar
operator
\eqref{eq:point-splitting-kg-operator} assigns to every globally hyperbolic
background \((M,g)\), every Hadamard state \(\omega\), and every test tensor
\(f^{\mu\nu}\in C_c^\infty(M,S^2TM)\) a number
\(\omega(T(f))\).  The prescription is required to satisfy the following
conditions.

\begin{enumerate}
\item It is obtained locally from
\begin{equation}
  \lim_{y\to x}\mathcal D_{\mu\nu}(x,y)
  \bigl(\omega_2(x,y)-H_{\epsilon,\mu}(x,y)\bigr)
  +C_{\mu\nu}[g](x),
  \label{eq:point-splitting-prescription-template}
\end{equation}
where \(\mathcal D_{\mu\nu}\) is the improved bidifferential operator
displayed in
\eqref{eq:point-split-stress-bidifferential-operator}, with
\(\eta=\eta_4=1/3\), and \(C_{\mu\nu}[g]\) is a state-independent local
curvature tensor.
\item It is locally covariant: if \(\chi:(M,g)\to(N,h)\) is an admissible
isometric embedding in the locally covariant category and
\(\omega_N\) is a Hadamard state on \(N\), then the pullback state on \(M\)
has stress-tensor expectation equal to the pullback of the \(N\)-expectation.
\item Its dependence on the state is exactly through the smooth biscalar
\(\omega_2-H_{\epsilon,\mu}\).  Equivalently, for two Hadamard states
\(\omega,\omega'\), the difference
\(\omega(T_{\mu\nu})-\omega'(T_{\mu\nu})\) is obtained by applying
\(\mathcal D_{\mu\nu}\) to \(\omega_2-\omega'_2\) and then taking the
diagonal limit.
\item It is conserved after the finite local tensor has been fixed:
\begin{equation}
  \nabla^\mu\omega(T_{\mu\nu})=0
  \label{eq:point-splitting-stress-conservation-condition}
\end{equation}
for all Hadamard states in the declared state space.
\item It has engineering dimension four under constant rescaling of the
metric and fields in the sense of
\eqref{eq:stress-tensor-engineering-dimensions}, and it is normalized so
that the Minkowski vacuum stress tensor vanishes when \(m=0\) and the finite
cosmological term is set to zero.
\end{enumerate}
\end{definition}

\subsection{The Conserved Bidifferential Prescription}

Let \(x\) and \(y\) lie in a common normal neighborhood, and let
\(g_\mu{}^{\mu'}(x,y)\) denote parallel transport of covectors from \(y\) to
\(x\).  Primed indices are differentiated at \(y\) and then transported to
\(x\).  For a smooth biscalar \(F\), put
\[
  F_{\rm s}(x,y)=\frac12\bigl(F(x,y)+F(y,x)\bigr),
  \qquad
  [A]_\Delta(x)=\lim_{y\to x}A(x,y).
\]
The polarized Hilbert tensor, including the classically vanishing
equation-of-motion improvement, is
\begin{equation}
\begin{aligned}
  \mathcal D^{(\eta)}_{\mu\nu}F
  ={}&
  \frac{1-2\xi}{2}
  \left(
    g_\nu{}^{\nu'}\nabla_\mu\nabla_{\nu'}
    +g_\mu{}^{\mu'}\nabla_\nu\nabla_{\mu'}
  \right)F_{\rm s} \\
  &+
  \left(2\xi-\frac12\right)
  g_{\mu\nu}g^{\rho\sigma'}\nabla_\rho\nabla_{\sigma'}F_{\rm s} \\
  &-\xi
  \left(
    \nabla_\mu\nabla_\nu
    +g_\mu{}^{\mu'}g_\nu{}^{\nu'}\nabla_{\mu'}\nabla_{\nu'}
  \right)F_{\rm s} \\
  &+\xi g_{\mu\nu}
  \left(\Box_x+\Box_y\right)F_{\rm s}
  +\xi G_{\mu\nu}F_{\rm s}
  -\frac12m^2g_{\mu\nu}F_{\rm s} \\
  &+\frac{\eta}{2}g_{\mu\nu}
  \left(P_{M,x}+P_{M,y}\right)F_{\rm s}.
\end{aligned}
  \label{eq:point-split-stress-bidifferential-operator}
\end{equation}
The last line is the point-split form of
\(\eta g_{\mu\nu}\phi P_M\phi\).  It vanishes on classical solutions, but it
does not vanish after subtracting a Hadamard parametrix because the parametrix
is only a local parametrix, not an exact bisolution.

\begin{quotedtheorem}[Moretti improved point-splitting conservation theorem]
\label{prop:parametrix-defect-conservation-repair}
Let \(D>2\), let \(P_M=-\Box_g+m^2+\xi R[g]\) with constant \(m^2\) and
constant \(\xi\), and let \(\omega\) be a Hadamard state on a smooth globally
hyperbolic spacetime.  Choose the local Hadamard parametrix in the same
normalization as Definition~\ref{def:local-hadamard-subtraction}, or
equivalently Moretti's \(Z_n\) normalization, with truncation order high
enough that the differentiated diagonal limits and one divergence exist
(\(n\ge3\) suffices in four dimensions).  For the parallel-transported
point-split Hilbert operator with the equation-of-motion ambiguity
\(\eta g_{\mu\nu}\phi P_M\phi\), the averaged point-split stress tensor built
from the nonambiguous local Hadamard parametrix is conserved if and only if
\[
  \eta_D=\frac{D}{2(D+2)} ,
  \qquad\text{so in four dimensions}\qquad
  \eta_4=\frac13 .
\]
In even dimensions the trace contains the local Hadamard/Seeley--DeWitt
coefficient, and changes of the Hadamard length scale add conserved local
tensors of the same engineering dimension.

This is an imported analytic theorem; the short transport calculation above
only fixes the local notation.  In the notation of this chapter, it imports the
coincidence-limit identities for \(P_{M,x}H\), \(P_{M,y}H\), and their first
diagonal derivatives: the two smooth parametrix defects in
\eqref{eq:naive-point-split-divergence-defect} combine to one locally
covariant covector \(\mathcal K_\nu[H]\), the divergence of the last line of
\eqref{eq:point-split-stress-bidifferential-operator} contributes
\(\eta\,\mathcal K_\nu[H]\), and the Hadamard coefficient theorem fixes the
naive coefficient to \(-\eta_D\).  We cite Moretti's improved point-splitting
construction for this coefficient identity and conservation theorem, with the
local Wick-product framework and finite renormalization setting supplied by
Hollands--Wald
\cite{Moretti2003StressEnergy,HollandsWald2001LocalWick,
HollandsWald2005Conservation,HollandsWald2015QFCS}.
\end{quotedtheorem}

\paragraph{Convention translation.}
The local algebra that is proved in this chapter is the convention
translation around the imported theorem.  For a polarized classical product
\(F(x,y)=\phi(x)\psi(y)\), the Hilbert identity gives
\[
  \nabla^\mu T_{\mu\nu}(\phi,\psi)
  =
  -\frac12\bigl((P_M\phi)\nabla_\nu\psi+(P_M\psi)\nabla_\nu\phi\bigr)
\]
when \(m^2\) and \(\xi\) are constant.  Rewriting this identity in
point-split notation gives, for the operator without the last line of
\eqref{eq:point-split-stress-bidifferential-operator},
\begin{equation}
  \nabla^\mu
  \left[\mathcal D_{\mu\nu}^{(0)}F\right]_\Delta
  =
  -\frac12
  \left[
    g_\nu{}^{\nu'}\nabla_{\nu'}P_{M,x}F_{\rm s}
    +\nabla_\nu P_{M,y}F_{\rm s}
  \right]_\Delta .
  \label{eq:naive-point-split-divergence-defect}
\end{equation}
Thus the naive point-split tensor is conserved on exact bisolutions, but it
has a local divergence defect on
\[
  F_\omega=\omega_2-H_{\epsilon,\mu},
  \qquad
  P_{M,x}F_\omega=-P_{M,x}H_{\epsilon,\mu},\qquad
  P_{M,y}F_\omega=-P_{M,y}H_{\epsilon,\mu}.
\]
Those smooth local defects are determined by the Hadamard recursion, but the
specific coincidence identity reducing their divergence to
\(-\eta_D\mathcal K_\nu[H]\) is imported from the Moretti theorem boundary
above.

Consequently, in four dimensions the prescription
\begin{equation}
  \omega(T_{\mu\nu}^{\rm ps})
  =
  \left[
    \mathcal D_{\mu\nu}^{(\eta_4)}
    \bigl(\omega_2-H_{\epsilon,\mu}\bigr)
  \right]_\Delta
  +C_{\mu\nu}^{\rm res}[g,m^2,\xi]
  \label{eq:conserved-point-split-stress-prescription}
\end{equation}
has the nonlocal state-dependent divergence removed:
\[
  \nabla^\mu
  \left[
    \mathcal D_{\mu\nu}^{(1/3)}
    \bigl(\omega_2-H_{\epsilon,\mu}\bigr)
  \right]_\Delta
  =
  0 .
\]
It is therefore conserved whenever the residual local tensor
\(C_{\mu\nu}^{\rm res}\) is conserved.  That residual tensor is precisely the
finite local freedom classified in
Proposition~\ref{prop:finite-local-freedom-scalar-stress-tensor}; adding it
preserves the Ward identity.

\begin{quotedtheorem}[Hollands--Wald finite-renormalization theorem]
\label{prop:finite-local-freedom-scalar-stress-tensor}
Let \(T_{\mu\nu}\) and \(T'_{\mu\nu}\) be two prescriptions satisfying
Definition~\ref{def:point-split-stress-tensor-prescription}.  Their
difference is state-independent by the state-dependence axiom.  The
Peetre--Slovak/Hollands--Wald locality, weak-regularity, and almost
homogeneous scaling theorem then reduces that difference to a finite natural
polynomial in the metric jet, \(m^2\), and the dimensionless coupling \(\xi\),
subject to conservation and the declared engineering dimension.  In four
dimensions, for fixed \(\xi\), the difference has the form
\begin{equation}
  T'_{\mu\nu}-T_{\mu\nu}
  =
  a_0m^4g_{\mu\nu}
  +a_1m^2G_{\mu\nu}
  +a_2I_{\mu\nu}
  +a_3J_{\mu\nu},
  \label{eq:stress-tensor-finite-local-freedom}
\end{equation}
with real constants \(a_i\).  Here
\begin{equation}
  G_{\mu\nu}=R_{\mu\nu}-\frac12Rg_{\mu\nu},
  \label{eq:einstein-tensor-definition-local-freedom}
\end{equation}
and
\begin{equation}
  I_{\mu\nu}
  =
  \frac{1}{\sqrt{|g|}}
  \frac{\delta}{\delta g^{\mu\nu}}
  \int \sqrt{|g|}\,R^2\,\dd^4x,
  \qquad
  J_{\mu\nu}
  =
  \frac{1}{\sqrt{|g|}}
  \frac{\delta}{\delta g^{\mu\nu}}
  \int \sqrt{|g|}\,R_{\alpha\beta}R^{\alpha\beta}\,\dd^4x,
  \label{eq:ij-curvature-tensors-definition}
\end{equation}
with compactly supported metric variations and boundary terms discarded.
This classification is a quoted structural theorem
\cite{HollandsWald2001LocalWick,HollandsWald2005Conservation,
HollandsWald2015QFCS}; the dimension count below is only the basis
enumeration after that finite-jet/naturality theorem has been invoked.
\end{quotedtheorem}

\paragraph{Basis enumeration after the theorem.}
The dimension-four local covariant candidates split by powers of \(m\).  The
only \(m^4\) tensor with two covariant indices and no derivatives is
\(m^4g_{\mu\nu}\), and it is conserved because \(m\) is constant and
\(\nabla g=0\).  At order \(m^2\) the possible curvature tensors are
\(m^2R_{\mu\nu}\) and \(m^2Rg_{\mu\nu}\).  Their divergence is
\[
  \nabla^\mu(\alpha R_{\mu\nu}+\beta Rg_{\mu\nu})
  =
  \left(\frac{\alpha}{2}+\beta\right)\nabla_\nu R,
\]
using the contracted Bianchi identity.  Conservation therefore forces
\(\beta=-\alpha/2\), giving the Einstein tensor \(G_{\mu\nu}\).

It remains to classify the curvature-squared part.  In four dimensions the
scalar local curvature densities of dimension four are spanned, modulo total
derivatives, by
\[
  R^2,\qquad R_{\alpha\beta}R^{\alpha\beta},\qquad
  R_{\alpha\beta\gamma\delta}R^{\alpha\beta\gamma\delta}.
\]
Their metric Euler derivatives are symmetric conserved tensors because the
corresponding integrated counterterms are diffeomorphism invariant.  The
dimensionless parameter \(\xi\) in the scalar stress tensor can change the
constants multiplying these local tensors but cannot create a new
dimension-four covariant tensor.  The four-dimensional Gauss--Bonnet combination
\[
  E_4
  =
  R_{\alpha\beta\gamma\delta}R^{\alpha\beta\gamma\delta}
  -4R_{\alpha\beta}R^{\alpha\beta}
  +R^2
\]
has vanishing metric variation on compactly supported variations in the
interior, since its integral is topological.  Hence the Euler derivative of
\(R_{\alpha\beta\gamma\delta}R^{\alpha\beta\gamma\delta}\) is a linear
combination of the Euler derivatives of \(R^2\) and
\(R_{\alpha\beta}R^{\alpha\beta}\).  These two remaining Euler derivatives are
precisely \(I_{\mu\nu}\) and \(J_{\mu\nu}\) in
\eqref{eq:ij-curvature-tensors-definition}.

No further independent local covariant conserved tensor of this dimension is
available inside the stated finite local freedom: terms with explicit
derivatives, such as
\(\nabla_\mu\nabla_\nu R\), \(g_{\mu\nu}\Box R\), and \(\Box R_{\mu\nu}\),
occur exactly in the Euler derivatives just described; total derivative
scalar counterterms give boundary contributions under compactly supported
metric variations; and tensors with more curvatures or more derivatives have
dimension larger than four.  Thus the quoted finite-renormalization theorem
specializes to the displayed four-parameter form.

\begin{proposition}[Curvature-squared Euler tensors]
\label{prop:curvature-squared-euler-tensors-point-splitting}
With the convention of
\eqref{eq:ij-curvature-tensors-definition}, and with the same inverse-metric
variation convention used in
Chapter~\ref{ch:curved-semiclassical-backreaction}, the
two independent curvature-squared Euler tensors are
\begin{equation}
\begin{aligned}
  I_{\mu\nu}
  &=
  2RR_{\mu\nu}
  -\frac12g_{\mu\nu}R^2
  +2g_{\mu\nu}\Box_gR
  -2\nabla_\mu\nabla_\nu R,\\
  J_{\mu\nu}
  &=
  2R_{\mu\alpha\nu\beta}R^{\alpha\beta}
  -\frac12g_{\mu\nu}R_{\alpha\beta}R^{\alpha\beta}
  -\nabla_\mu\nabla_\nu R
  +\Box_gR_{\mu\nu}
  +\frac12g_{\mu\nu}\Box_gR .
\end{aligned}
  \label{eq:curvature-squared-euler-tensor-formulas}
\end{equation}
Thus \(J_{\mu\nu}=H^{(2)}_{\mu\nu}\) in
Chapter~\ref{ch:curved-semiclassical-backreaction};
the semiclassical chapter is not introducing an independent Ricci-squared sign
coordinate.
They are conserved,
\begin{equation}
  \nabla^\mu I_{\mu\nu}=0,\qquad
  \nabla^\mu J_{\mu\nu}=0,
  \label{eq:curvature-squared-euler-tensor-conservation}
\end{equation}
and in four dimensions their traces are
\begin{equation}
  g^{\mu\nu}I_{\mu\nu}=6\Box_gR,\qquad
  g^{\mu\nu}J_{\mu\nu}=2\Box_gR .
  \label{eq:curvature-squared-euler-tensor-traces}
\end{equation}
\end{proposition}

\begin{proof}
For the \(R^2\) functional, use
\begin{equation}
  \delta\sqrt{|g|}
  =
  -\frac12\sqrt{|g|}\,g_{\mu\nu}\delta g^{\mu\nu},
  \qquad
  \delta R
  =
  R_{\mu\nu}\delta g^{\mu\nu}
  +g_{\mu\nu}\Box_g\delta g^{\mu\nu}
  -\nabla_\mu\nabla_\nu\delta g^{\mu\nu}.
  \label{eq:variation-scalar-curvature-inverse-metric}
\end{equation}
After two integrations by parts with compactly supported metric variation,
\[
  \delta\int\sqrt{|g|}R^2
  =
  \int\sqrt{|g|}
  \left(
    2RR_{\mu\nu}
    -\frac12g_{\mu\nu}R^2
    +2g_{\mu\nu}\Box_gR
    -2\nabla_\mu\nabla_\nu R
  \right)\delta g^{\mu\nu},
\]
which is the first line of
\eqref{eq:curvature-squared-euler-tensor-formulas}.  For
\(\int\sqrt{|g|}R_{\alpha\beta}R^{\alpha\beta}\), the variation starts from
\[
  \delta\!\left(\sqrt{|g|}R_{\alpha\beta}R^{\alpha\beta}\right)
  =
  \sqrt{|g|}
  \left(
    -\frac12 g_{\mu\nu}R_{\alpha\beta}R^{\alpha\beta}
    +2R_{\mu\alpha}R_\nu{}^\alpha
  \right)\delta g^{\mu\nu}
  +2\sqrt{|g|}R^{\mu\nu}\delta R_{\mu\nu},
\]
using the covariant-metric Palatini identity
\begin{equation}
  \delta R_{\mu\nu}
  =
  \frac12\left(
    -\Box_g\delta g_{\mu\nu}
    -\nabla_\mu\nabla_\nu(g^{\alpha\beta}\delta g_{\alpha\beta})
    +\nabla^\alpha\nabla_\mu\delta g_{\alpha\nu}
    +\nabla^\alpha\nabla_\nu\delta g_{\alpha\mu}
  \right),
  \label{eq:variation-ricci-covariant-metric}
\end{equation}
then convert \(\delta g_{\mu\nu}\) to
\(-g_{\mu\alpha}g_{\nu\beta}\delta g^{\alpha\beta}\), integrate derivatives
off \(\delta g^{\mu\nu}\), and commute covariant derivatives on
\(R_{\mu\nu}\).  The minus sign in this conversion fixes the derivative
signs.  The integrated Palatini term contributes
\[
  -\nabla_\mu\nabla_\nu R+\Box_gR_{\mu\nu}
  +\frac12g_{\mu\nu}\Box_gR ,
\]
while the commutator terms combine with the explicit metric variation of
\(R^{\alpha\beta}\) into
\(2R_{\mu\alpha\nu\beta}R^{\alpha\beta}\).  This gives the second line.
These are ordinary variational identities for compactly supported variations;
no field equation is used.

Conservation follows from diffeomorphism invariance of the two functionals.
For a compactly supported infinitesimal diffeomorphism generated by
\(\zeta^\mu\),
\[
  \delta_\zeta g^{\mu\nu}
  =
  -\nabla^\mu\zeta^\nu-\nabla^\nu\zeta^\mu .
\]
If \(E_{\mu\nu}\) denotes either variational derivative, then
\[
  0=\delta_\zeta\mathcal F
  =
  -2\int\sqrt{|g|}\,E_{\mu\nu}\nabla^\mu\zeta^\nu
  =
  2\int\sqrt{|g|}\,(\nabla^\mu E_{\mu\nu})\zeta^\nu ,
\]
so \(\nabla^\mu E_{\mu\nu}=0\) distributionally, hence smoothly.  Taking
the trace of the two displayed formulas and using
\(g^{\mu\nu}\Box_gR_{\mu\nu}=\Box_gR\) gives
\[
  g^{\mu\nu}I_{\mu\nu}
  =
  (2-\tfrac D2)R^2+(2D-2)\Box_gR,
  \qquad
  g^{\mu\nu}J_{\mu\nu}
  =
  (2-\tfrac D2)R_{\alpha\beta}R^{\alpha\beta}
  +\frac D2\Box_gR .
\]
Setting \(D=4\) gives
\eqref{eq:curvature-squared-euler-tensor-traces}.
\end{proof}

\section{Using the Bidifferential Operator}

Equation~\eqref{eq:point-split-stress-bidifferential-operator} is the
operator used in the examples below.  If the argument is a difference of two
Hadamard two-point functions, it is a smooth bisolution and the
equation-of-motion line contributes zero.  If the argument is
\(\omega_2-H_{\epsilon,\mu}\), the same line is essential: it cancels the
local parametrix defect described in
Proposition~\ref{prop:parametrix-defect-conservation-repair}.  The remaining
choice is the conserved residual local tensor in
\eqref{eq:conserved-point-split-stress-prescription}, whose finite freedom is
classified by
Proposition~\ref{prop:finite-local-freedom-scalar-stress-tensor}.

\section{Worked Flat-Space Computation}

We now compute a simple case explicitly.  Let
\((M,\eta)\) be four-dimensional Minkowski spacetime with
\(\eta={\rm diag}(-1,1,1,1)\), and take the massless minimally coupled free
scalar.  The vacuum two-point distribution is
\[
  W_0(x,x')
  =
  \lim_{\epsilon\downarrow0}
  \frac{1}{4\pi^2}
  \frac{1}{
  |\mathbf x-\mathbf x'|^2-(t-t'-\ii\epsilon)^2}.
\]
In flat space the local Hadamard parametrix with the same \(i\epsilon\)
prescription is \(H_0=W_0\).  Therefore the renormalized vacuum stress
tensor obtained by point splitting is zero:
\[
  \langle T_{\mu\nu}\rangle_{0,{\rm ren}}
  =
  \lim_{x'\to x}
  \mathcal D_{\mu\nu}(x,x')\bigl(W_0(x,x')-H_0(x,x')\bigr)
  =
  0 ,
\]
if the finite flat-space cosmological-constant counterterm is set to zero.
This is the explicit subtraction of the state-independent short-distance
singularity.

For the thermal KMS state at inverse temperature \(\beta\) with respect to
the inertial time \(t\), the difference from the vacuum two-point function is
smooth:
\[
  \Delta W_\beta(x,x')
  =
  W_\beta(x,x')-W_0(x,x')
\]
\[
  =
  \int \frac{d^3k}{(2\pi)^3}\,
  \frac{n_\beta(\omega)}{2\omega}
  \left(
  e^{-\ii\omega(t-t')+\ii\mathbf k\cdot(\mathbf x-\mathbf x')}
  +
  e^{+\ii\omega(t-t')-\ii\mathbf k\cdot(\mathbf x-\mathbf x')}
  \right),
  \qquad
  \omega=|\mathbf k|,
\]
where \(n_\beta(\omega)=(e^{\beta\omega}-1)^{-1}\).  Since
\(\Delta W_\beta\) is smooth, the point-split stress tensor is obtained by
applying the polarized classical expression directly to this remainder.
For the massless minimal tensor
\[
  T_{00}
  =
  \frac{1}{2}\bigl((\partial_t\phi)^2+|\nabla\phi|^2\bigr),
\]
the bidifferential operator is
\[
  \mathcal D_{00}
  =
  \frac{1}{2}\left(
  \partial_t\partial_{t'}
  +\sum_{i=1}^3\partial_i\partial_{i'}
  \right).
\]
On either plane wave in \(\Delta W_\beta\),
\[
  \mathcal D_{00}
  e^{-\ii\omega(t-t')+\ii\mathbf k\cdot(\mathbf x-\mathbf x')}
  =
  \omega^2
  e^{-\ii\omega(t-t')+\ii\mathbf k\cdot(\mathbf x-\mathbf x')}.
\]
Taking the diagonal limit gives
\[
  \rho_\beta
  =
  \langle T_{00}\rangle_{\beta,{\rm ren}}
  =
  \int \frac{d^3k}{(2\pi)^3}\,\omega\,n_\beta(\omega)
  =
  \frac{1}{2\pi^2}\int_0^\infty
  \frac{k^3\,dk}{e^{\beta k}-1}.
\]
The change of variables \(x=\beta k\) and the expansion
\((e^x-1)^{-1}=\sum_{n\geq1}e^{-nx}\) give
\[
  \int_0^\infty \frac{x^3\,dx}{e^x-1}
  =
  \sum_{n\geq1}\int_0^\infty x^3e^{-nx}\,dx
  =
  6\sum_{n\geq1}\frac{1}{n^4}
  =
  \frac{\pi^4}{15}.
\]
Thus
\[
  \rho_\beta=\frac{\pi^2}{30\beta^4}.
\]
For the spatial components,
\[
  T_{ij}
  =
  \partial_i\phi\,\partial_j\phi
  +\frac{1}{2}\delta_{ij}
  \bigl((\partial_t\phi)^2-|\nabla\phi|^2\bigr),
\]
so
\[
  \mathcal D_{ij}
  =
  \partial_i\partial_{j'}
  +\frac{1}{2}\delta_{ij}
  \left(
  \partial_t\partial_{t'}
  -\sum_{\ell=1}^3\partial_\ell\partial_{\ell'}
  \right).
\]
For massless modes the second term vanishes on shell because
\(\omega^2=|\mathbf k|^2\).  Rotational invariance of the momentum integral
gives
\[
  \int d\Omega\, k_i k_j=\frac{4\pi}{3}\delta_{ij}k^2,
\]
and therefore
\[
  \langle T_{ij}\rangle_{\beta,{\rm ren}}
  =
  \frac{1}{3}\rho_\beta\,\delta_{ij}.
\]
The point-split computation thus yields
\[
  \langle T_{\mu\nu}\rangle_{\beta,{\rm ren}}
  =
  \frac{\pi^2}{90\beta^4}
  \begin{pmatrix}
  3&0&0&0\\
  0&1&0&0\\
  0&0&1&0\\
  0&0&0&1
  \end{pmatrix}_{\mu\nu},
\]
in the inertial frame.  Its trace is
\(-\rho_\beta+3p_\beta=0\), as required for the massless conformal equation
of state in flat spacetime.

\section{Constant-Curvature Diagnostic}

A de Sitter-invariant Hadamard state of a free scalar has a stress-tensor
expectation value constrained by symmetry to the form
\[
  \langle T_{\mu\nu}\rangle_{\rm dS}=C\,g_{\mu\nu}.
\]
For the massless conformally coupled real scalar in four dimensions, the
state-dependent classical trace vanishes.  With the trace-anomaly convention
of Chapter~\ref{ch:curved-trace-anomalies}, the real scalar has
\(a=1/360\), \(c=1/120\), and
\[
  \langle T^\mu{}_\mu\rangle
  =
  \frac{1}{16\pi^2}\left(cW_{\mu\nu\rho\sigma}W^{\mu\nu\rho\sigma}
  -aE_4+b\Box R\right).
\]
On four-dimensional de Sitter with Hubble parameter \(H\),
\[
  W_{\mu\nu\rho\sigma}=0,\qquad
  \Box R=0,\qquad
  R=12H^2,
\]
\[
  R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma}=24H^4,\quad
  R_{\mu\nu}R^{\mu\nu}=36H^4,\quad
  E_4=24H^4 .
\]
Consequently
\[
  4C=\langle T^\mu{}_\mu\rangle
  =
  -\frac{H^4}{240\pi^2},
  \qquad
  C=-\frac{H^4}{960\pi^2}.
\]
This calculation isolates the part fixed by local covariance, de Sitter
invariance, and the anomaly normalization.  In the massless conformal case,
the classified four-parameter freedom does not shift this constant once the
flat cosmological term has been set to zero: the \(m^4g_{\mu\nu}\) and
\(m^2G_{\mu\nu}\) terms vanish, while \(I_{\mu\nu}\) and \(J_{\mu\nu}\) vanish
on four-dimensional constant-curvature metrics.  Massive fields can shift
\(C\) through the \(m\)-dependent local tensors, and any further shift would
come from leaving the stated locally covariant, weakly regular,
almost-homogeneous scaling class.

\section{Conservation and Trace}

\begin{definition}[Conserved stress tensor]
\label{def:conserved-renormalized-stress-tensor-curved}
Within a locally covariant point-splitting scheme, a renormalized stress
tensor is called conserved when its finite local curvature terms have been
chosen so that
\[
  \nabla^\mu\omega(T_{\mu\nu})=0
\]
for all Hadamard states in the declared class.
\end{definition}

The trace convention is fixed by the classical Hilbert tensor, not by the
anomaly example.  In \(D\) dimensions,
\begin{equation}
\begin{aligned}
  T^\mu{}_\mu
  ={}&
  \left(1-\frac D2+2\xi(D-1)\right)(\nabla\phi)^2
  +2\xi(D-1)\phi\Box_g\phi \\
  &-\frac D2m^2\phi^2
  +\xi\left(1-\frac D2\right)R\phi^2 .
\end{aligned}
  \label{eq:scalar-hilbert-trace-before-eom}
\end{equation}
Using \(P_M\phi=0\), equivalently
\(\Box_g\phi=(m^2+\xi R)\phi\), and
\(\Box_g\phi^2=2(\nabla\phi)^2+2\phi\Box_g\phi\), this becomes
\begin{equation}
  T^\mu{}_\mu
  =
  -m^2\phi^2
  +(D-1)(\xi-\xi_c)\Box_g\phi^2,
  \qquad
  \xi_c=\frac{D-2}{4(D-1)} .
  \label{eq:scalar-classical-trace-corrected}
\end{equation}
The corresponding renormalized composite identity is therefore
\begin{equation}
  \omega(T^\mu{}_\mu)
  =
  -m^2\,\omega(:\!\Phi^2\!:)
  +(D-1)(\xi-\xi_c)\,
    \omega(\Box_g:\!\Phi^2\!:)
  +\mathcal A[g,m,\xi],
  \label{eq:renormalized-scalar-trace-corrected}
\end{equation}
where \(\mathcal A\) is a local curvature scalar determined up to the same
finite renormalization freedom.  It includes the local remnants of the
parametrix equation-of-motion defect, contact terms in the composite
equation of motion, and the trace variation of the chosen finite
counterterms.  In the massless conformally coupled case, the state-dependent
classical trace terms vanish and the remaining local term is the trace
anomaly.

\section{Relation to Flat-Space Composite Operators}

Point splitting is an operator-definition scheme.  It differs from defining a
regularized composite operator by inserting a classical functional into a
regulated path integral and then renormalizing.  The two schemes can be
compared by finite local counterterms when both are defined on the same
background and state class.  This comparison is the curved-background version
of the operator-coordinate changes developed in the renormalization volume.

\begin{openproblem}[Interacting stress tensors beyond formal perturbation]
\label{op:interacting-stress-tensors-beyond-formal-perturbation}
Construct stress tensors for interacting curved-spacetime QFTs with
nonperturbative control of domains, local covariance, conservation, trace
anomalies, and state dependence.  A solution should identify the exact
relation between locally covariant composite fields and flat-space
renormalized operator schemes.
\end{openproblem}
