\chapter{Eta Invariants and Global Anomalies}
\label{ch:curved-eta-invariants-global-anomalies}

\ConventionsEuclidean

Chapter~\ref{ch:curved-background-gauge-fields-index-theory} treated the
local index density and the curvature of the determinant line.  This chapter
develops the complementary holonomy problem.  A local anomaly is detected by
infinitesimal motion in background-field space; a global anomaly is detected
by transporting the fermion functional integral around a closed loop of
backgrounds.  For free fermions the transport is computed by eta invariants
of Dirac operators on the associated mapping tori.
The source traces for the global-analysis inputs are the APS spectral
asymmetry papers, the Bismut--Freed holonomy theorem, Dai--Freed determinant
lines, Freed's graded determinant-line convention, and Witten's \(SU(2)\)
anomaly
\cite{APS1975I,APS1975II,APS1976III,BismutFreed1986,DaiFreed1994,
Freed1995DeterminantRevisited,Witten1982}.

\section{Self-Adjoint Dirac Operators and Eta Functions}

Let \(Y\) be a closed oriented Riemannian spin manifold of odd dimension
\(2n+1\).  Let \(E\to Y\) be a Hermitian vector bundle with unitary connection
\(\nabla^E\).  The coupled Dirac operator
\[
  B_Y
  :
  C^\infty(Y,S_Y\otimes E)\longrightarrow
  C^\infty(Y,S_Y\otimes E)
\]
is self-adjoint on \(L^2(Y,S_Y\otimes E)\) when the Clifford action is chosen
skew-Hermitian in Euclidean signature.  Its spectrum is real, discrete, and
has finite-dimensional eigenspaces.  The eta function is initially defined by
\begin{equation}
  \eta_{B_Y}(s)
  =
  \sum_{\lambda\in\operatorname{Spec}(B_Y)\setminus\{0\}}
  \operatorname{sign}(\lambda)|\lambda|^{-s},
  \qquad
  \operatorname{Re}s\gg1 .
  \label{eq:eta-function-definition}
\end{equation}
For a Dirac-type operator the heat-kernel expansion gives a meromorphic
continuation to \(s\in\mathbb C\), regular at \(s=0\).  We define
\begin{equation}
  h(B_Y)=\dim\ker B_Y,
  \qquad
  \xi(B_Y)=\frac{\eta_{B_Y}(0)+h(B_Y)}2 .
  \label{eq:reduced-eta-invariant}
\end{equation}
The real number \(\xi(B_Y)\) depends on the metric and connection, but
\(\exp(2\pi\ii\xi(B_Y))\) is the natural phase variable: when eigenvalues
cross through zero, \(\xi\) changes by an integer.

\begin{remark}[Integer jump at a simple crossing]
\label{rem:eta-integer-jump}
Let \(B_t\), \(t\in(-\varepsilon,\varepsilon)\), be a smooth one-parameter
family of self-adjoint Dirac-type operators.  Suppose exactly one simple
eigenvalue \(\lambda(t)\) crosses zero, with
\(\lambda(t)<0\) for \(t<0\) and \(\lambda(t)>0\) for \(t>0\), and all other
eigenvalues stay away from zero.  Then the contribution of this eigenvalue to
\(\xi(B_t)\) jumps by \(+1\).  A crossing in the opposite direction gives a
jump by \(-1\).
Away from the crossing the eigenvalue contributes
\(\operatorname{sign}(\lambda)/2\) to
\((\eta(0)+h)/2\).  Just before the crossing this contribution is
\(-1/2\), at the crossing it is \(1/2\) because \(h=1\), and just after the
crossing it is \(1/2\).  Thus the one-sided jump from \(t<0\) to \(t>0\) is
\(+1\).  If the eigenvalue crosses from positive to negative, the
contribution changes from \(+1/2\) before the crossing to \(-1/2\) after the
crossing, again with \(h=1\) at the crossing instant, so the one-sided jump is
\(-1\).
\end{remark}

\begin{definition}[Spectral flow]
For a smooth path \(B_t\), \(0\leq t\leq1\), whose endpoints are invertible,
the spectral flow \(\operatorname{sf}(B_t)\) is the signed number of
eigenvalues crossing zero from negative to positive, counted with
multiplicity.  If the endpoints have kernels, one fixes a small perturbation
or uses the APS convention below; the resulting quantity is well-defined
modulo the endpoint-kernel bookkeeping appearing in \(\xi\).
\end{definition}

Remark~\ref{rem:eta-integer-jump} is the elementary reason that global
anomalies are naturally phases rather than real-valued functions of
\(\eta(0)\) alone.  The smooth part of \(\xi\) is local; the integer jumps
record spectral flow.

\section{The APS Boundary Formula}

Let \(X\) be a compact oriented Riemannian spin manifold of even dimension
\(2n+2\), with boundary \(\partial X=Y\).  Assume that, on a collar
\((-\epsilon,0]\times Y\), the metric and the connection have product form.
Let \(u\) be the inward normal coordinate.  In this collar the chiral Dirac
operator has the form
\begin{equation}
  D_X^+
  =
  c(\nu)\bigl(\partial_u+B_Y\bigr)
  \quad\hbox{up to the standard chiral identification,}
  \label{eq:aps-collar-dirac-form}
\end{equation}
where \(\nu\) is the outward unit normal and \(B_Y\) is the self-adjoint
tangential Dirac operator on \(Y\).

The APS boundary condition keeps the boundary values of spinors in the
negative spectral subspace of \(B_Y\):
\begin{equation}
  P_{\geq0}\bigl(\psi|_Y\bigr)=0,
  \label{eq:aps-boundary-condition}
\end{equation}
where \(P_{\geq0}\) is the orthogonal projection onto the span of eigenvectors
with eigenvalue \(\geq0\).  Including the zero eigenspace in \(P_{\geq0}\) is
the convention paired with the reduced invariant \(\xi=(\eta+h)/2\).

\paragraph{Half-cylinder sign check.}
The sign in \eqref{eq:aps-boundary-condition} is fixed by the product collar,
not by an arbitrary convention.  Replace the collar by the infinite inward
cylinder \((-\infty,0]\times Y\) with the same model operator
\[
  c(\nu)(\partial_u+B_Y).
\]
If \(B_Y\phi_\lambda=\lambda\phi_\lambda\) and
\(\psi(u,y)=f_\lambda(u)\phi_\lambda(y)\), the equation
\((\partial_u+B_Y)\psi=0\) becomes
\[
  \frac{d f_\lambda}{du}+\lambda f_\lambda=0,
  \qquad
  f_\lambda(u)=e^{-\lambda u}f_\lambda(0).
\]
Since the completed cylinder runs to \(u=-\infty\), this mode is square
integrable exactly when \(\lambda<0\).  Thus the boundary values of \(L^2\)
solutions on the completed model lie in the negative spectral subspace, while
zero modes are not square-integrable on the infinite cylinder and are
therefore placed in the forbidden subspace \(P_{\ge0}\).  On an oppositely
oriented boundary component the tangential operator is \(-B_Y\), so the same
mode calculation reverses the spectral sign.  This elementary model is not the
APS Fredholm theorem; it fixes the boundary-orientation convention used in the
eta and anomaly formulae below.

\begin{quotedtheorem}[Atiyah--Patodi--Singer index formula]
\label{qthm:aps-index-formula}
\emph{Source trace.}  This is the APS boundary index theorem in the reduced
eta convention used in this chapter
\cite{APS1975I,APS1975II,APS1976III}.
For the product-collar data above, the chiral Dirac operator with APS
boundary condition is Fredholm and
\begin{equation}
  \operatorname{Ind}(D_X^{+,\mathrm{APS}})
  =
  \int_X
  \bigl[\widehat A(TX)\operatorname{ch}(E)\bigr]_{2n+2}
  -
  \xi(B_Y).
  \label{eq:aps-index-formula}
\end{equation}
For non-product data one first deforms to product form near the boundary or
adds the corresponding local transgression term.  The sum of bulk density,
local boundary transgression, and eta term is the invariant integer.
\end{quotedtheorem}

\paragraph{Analytic mechanism behind the boundary term.}
The theorem is used below through three precise pieces of analytic structure.
First, product form near the boundary identifies the normal part of the Dirac
operator with the first-order operator
\eqref{eq:aps-collar-dirac-form}.  The spectral projection
\(P_{\geq0}\) then makes the boundary problem elliptic in the
pseudodifferential sense: boundary values in the forbidden spectral subspace
are removed, while the remaining components give a Fredholm operator.  Second,
the small-time heat-kernel expansion of the APS problem has the same interior
coefficient \([\widehat A(TX)\operatorname{ch}(E)]_{2n+2}\) as the closed
index theorem.  This is the local part of the formula and is the source of the
usual anomaly polynomial.  Third, the APS boundary condition is nonlocal along
\(Y\), and its nonlocal contribution is exactly the reduced spectral
asymmetry \(\xi(B_Y)\).  The integer on the left side of
\eqref{eq:aps-index-formula} is the compatibility condition tying these local
and nonlocal terms together.

In this chapter the APS theorem functions as the global-analysis bridge
between two objects already defined above: the local index density on \(X\) and
the spectral invariant of the boundary Dirac operator \(B_Y\).  The QFT
statement extracted from it is the following: local anomaly density determines
the infinitesimal variation of the fermion phase, while eta invariants retain
the integral spectral-flow data that can survive after the local curvature has
vanished.

\paragraph{Cylinder variation of the reduced eta invariant.}
Let \(B_t\), \(0\leq t\leq1\), be a smooth path of odd-dimensional Dirac
operators obtained from a smooth path of metrics and unitary connections on a
fixed closed odd-dimensional spin manifold \(Y\).  Let
\[
  X=[0,1]\times Y
\]
carry the interpolating product-near-boundary data.  Then
\begin{equation}
  \xi(B_1)-\xi(B_0)
  \equiv
  \int_X
  \bigl[\widehat A(TX)\operatorname{ch}(E)\bigr]_{2n+2}
  \pmod{\mathbb Z}.
  \label{eq:eta-cylinder-variation}
\end{equation}
If the path has no zero crossings and the endpoint kernels have constant
dimension, differentiating this formula gives the usual local descent
variation of \(\xi\).
This is the cylinder specialization of APS.  The boundary of
\(X=[0,1]\times Y\) is
\[
  \partial X=Y_1\sqcup(-Y_0).
\]
The tangential operator on \(Y_1\) is \(B_1\), while the tangential operator on
the oppositely oriented component \(-Y_0\) is \(-B_0\).  Applying
\eqref{eq:aps-index-formula} to \(X\) gives
\[
  \operatorname{Ind}(D_X^{+,\mathrm{APS}})
  =
  \int_X
  \bigl[\widehat A(TX)\operatorname{ch}(E)\bigr]_{2n+2}
  -
  \xi(B_1)-\xi(-B_0).
\]
Since \(\eta_{-B_0}(0)=-\eta_{B_0}(0)\) and
\(h(-B_0)=h(B_0)\), one has
\[
  \xi(-B_0)
  =
  -\xi(B_0)+h(B_0).
\]
The Fredholm index and \(h(B_0)\) are integers.  Moving them to the other
side gives \eqref{eq:eta-cylinder-variation} modulo \(\mathbb Z\).

\paragraph{Exact cylinder bookkeeping.}
The preceding congruence is often the only part of the APS cylinder formula
that survives in anomaly phases, but the exact integer identity is what fixes
the sign conventions.  With
\[
  I_X(E)
  =
  \int_X
  \bigl[\widehat A(TX)\operatorname{ch}(E)\bigr]_{2n+2},
\]
the APS formula on \(\partial X=Y_1\sqcup(-Y_0)\) gives
\begin{equation}
  \operatorname{Ind}(D_X^{+,\mathrm{APS}})
  =
  I_X(E)-\xi(B_1)+\xi(B_0)-h(B_0).
  \label{eq:aps-cylinder-exact-bookkeeping}
\end{equation}
Equivalently,
\begin{equation}
  \xi(B_1)-\xi(B_0)
  =
  I_X(E)-h(B_0)-\operatorname{Ind}(D_X^{+,\mathrm{APS}}).
  \label{eq:aps-cylinder-xi-exact-bookkeeping}
\end{equation}
Thus the endpoint-kernel convention is not an afterthought: including the
zero eigenspace in \(P_{\ge0}\) at the incoming boundary contributes the
integer \(h(B_0)\) in the exact formula and then disappears only after
passing to \(\mathbb R/\mathbb Z\).  When \(B_0\) and \(B_1\) are invertible,
\eqref{eq:aps-cylinder-exact-bookkeeping} becomes
\[
  \operatorname{Ind}(D_X^{+,\mathrm{APS}})
  =
  I_X(E)-\bigl(\xi(B_1)-\xi(B_0)\bigr).
\]
If the path has only transverse simple crossings, the integer jump part of
\(\xi(B_1)-\xi(B_0)\) is exactly the spectral flow
\(\operatorname{sf}(B_t)\), by Remark~\ref{rem:eta-integer-jump}.  The
remaining smooth variation is represented by the local transgression term
whose cylinder integral is \(I_X(E)\); the APS index supplies the integer
needed to make the exact equality hold.
Consequently, in the orientation and APS boundary convention of
\eqref{eq:aps-collar-dirac-form}--\eqref{eq:aps-boundary-condition}, the
pure crossing contribution satisfies
\[
  \operatorname{Ind}(D_X^{+,\mathrm{APS}})_{\mathrm{cross}}
  =
  -\,\operatorname{sf}(B_t).
\]
This minus sign is a boundary-orientation convention, not a physical phase:
the exponentiated eta phase only remembers the congruence
\eqref{eq:eta-cylinder-variation}, while the exact identity above records the
Fredholm representative used to compute it.

Figure~\ref{fig:eta-cylinder-mapping-torus} records the cylinder/mapping-torus
geometry used in the eta-variation formula: interpolating between boundary
operators converts eta change into an APS index modulo integers.

\begin{figure}[t]
\centering
\begin{tikzpicture}[>=Stealth,scale=0.95]
  \draw[thick] (0,0) rectangle (4.2,2.2);
  \draw[->,thick] (0.45,1.1) -- (3.75,1.1);
  \node at (0.2,1.1) [left] {\(Y_0\)};
  \node at (4.4,1.1) [right] {\(Y_1\)};
  \node at (2.1,2.45) {\([0,1]\times Y\)};
  \node at (2.1,0.75) {\(B_t\)};
  \draw[rounded corners,thick] (6.0,0.15) rectangle (9.9,2.05);
  \draw[->,thick] (6.5,1.1) arc[start angle=180,end angle=-180,x radius=1.45,y radius=0.72];
  \node at (7.95,2.45) {mapping torus \(Y_\gamma\)};
  \node at (7.95,0.65) {\((x,1)\sim\gamma(x,0)\)};
\end{tikzpicture}
\caption{The APS cylinder computes the local change of \(\xi\) modulo
integers.  Identifying the two ends by a closed loop of backgrounds gives the
mapping torus whose eta invariant computes determinant-line holonomy.}
\label{fig:eta-cylinder-mapping-torus}
\end{figure}

\section{Determinant-Line Holonomy}

Let \(M\) be a closed even-dimensional spin manifold, and let
\(\mathcal B\) be a smooth parameter space of background metrics and
connections, before quotienting by large gauge transformations.  A family of
chiral Dirac operators
\[
  D_b^+:\Gamma(M,S^+\otimes E_b)\longrightarrow
  \Gamma(M,S^-\otimes E_b),
  \qquad b\in\mathcal B,
\]
defines the determinant superline
\begin{equation}
  \operatorname{Det}D_b
  =
  \bigwedge\nolimits^{\mathrm{top}}\ker D_b^+
  \otimes
  \left(\bigwedge\nolimits^{\mathrm{top}}\ker D_b^-\right)^* .
  \label{eq:det-line-fiber}
\end{equation}
The underlying vector space is one-dimensional, but in the Dai--Freed/Freed
formalism it is a complex \(\mathbb Z_2\)-graded line.  Its parity is
\begin{equation}
  p(\operatorname{Det}D_b)
  \equiv
  \operatorname{Ind}(D_b^+)
  \equiv
  \dim\ker D_b^+ - \dim\ker D_b^-
  \pmod 2 .
  \label{eq:det-superline-parity}
\end{equation}
We write \(L^\vee\) for the graded dual.  As an ungraded line this is the
usual inverse line, but in a gluing formula it retains its parity and the
graded tensor category supplies Koszul signs.
Where kernels jump this formula is not a literal vector bundle chart, but the
Quillen determinant construction glues these local descriptions to a smooth
complex superline bundle over \(\mathcal B\).  The regulated chiral fermion
functional integral is a section of this superline, not a function, until a
trivialization is chosen.

Let \(\pi:\mathcal M\to\mathcal B\) denote the resulting family of closed
even-dimensional spin manifolds with vertical metric, vertical spin
structure, coupled bundle, and a chosen horizontal distribution.  In the
present fixed-manifold background-field setting this family is often the
product \(\mathcal B\times M\), but the horizontal distribution is still the
geometric datum used to form the odd-dimensional Dirac operator along a path
in \(\mathcal B\).

Let \(\gamma:S^1\to\mathcal B\) be a loop of backgrounds.  Pulling the family
back to the circle and gluing the ends gives the mapping torus
\begin{equation}
  Y_\gamma
  =
  \gamma^*\mathcal M
  =
  ([0,1]\times M)/\bigl((1,x)\sim(0,x)\hbox{ using }\gamma\bigr).
  \label{eq:mapping-torus-background-loop}
\end{equation}
Choose a spin structure \(\sigma\) on the base circle.  Together with the
vertical spin structure on \(T(Y_\gamma/S^1)\), it induces a spin structure
on \(Y_\gamma\).  For \(\varepsilon>0\) set
\begin{equation}
  g_{\gamma,\varepsilon}
  =
  \varepsilon^{-2}d\theta^2
  \oplus
  g^{T(Y_\gamma/S^1)},
  \label{eq:bismut-freed-adiabatic-metric}
\end{equation}
using the horizontal/vertical splitting.  Let
\(B_{\gamma,\varepsilon}^{\sigma}\) denote the resulting odd Dirac operator
on \(Y_\gamma\).  The unscaled eta invariant of an arbitrary metric on
\(Y_\gamma\) is not the determinant-line holonomy; the holonomy is recovered
only after the adiabatic limit \(\varepsilon\to0\).

\begin{quotedtheorem}[Bismut--Freed holonomy formula]
\label{qthm:bismut-freed-holonomy}
\emph{Source trace.}  The determinant-line connection and eta-holonomy
formula are the Bismut--Freed holonomy theorem in the Dirac-family case,
with Freed's adiabatic-limit and circle-spin conventions made explicit
\cite{BismutFreed1986,Freed1995DeterminantRevisited}.
With the determinant-line and chirality conventions of
\eqref{eq:det-line-fiber}, this is the dual of Freed's inverse determinant
line \(L=(\operatorname{Det}D)^\vee\).  Dualizing the connection inverts its
\(U(1)\) holonomy, so the eta exponent changes sign; the circle-spin factor
\((-1)^{\operatorname{Ind}(D^+)}\) is unchanged because it is self-inverse.
Thus the natural Bismut--Freed connection on \(\operatorname{Det}D\) has
holonomy
\begin{equation}
  \operatorname{Hol}_{\gamma,\sigma}(\operatorname{Det}D)
  =
  \begin{cases}
  \displaystyle
  \lim_{\varepsilon\to0}
  \exp\!\left(-2\pi\ii\,\xi(B_{\gamma,\varepsilon}^{\sigma_{\rm b}})\right),
  & \sigma=\sigma_{\rm b},\\[3mm]
  \displaystyle
  (-1)^{\operatorname{Ind}(D^+)}
  \lim_{\varepsilon\to0}
  \exp\!\left(-2\pi\ii\,\xi(B_{\gamma,\varepsilon}^{\sigma_{\rm nb}})\right),
  & \sigma=\sigma_{\rm nb},
  \end{cases}
  \label{eq:bismut-freed-holonomy}
\end{equation}
where \(\sigma_{\rm b}\) is the bounding spin structure on the mapping
circle and \(\sigma_{\rm nb}\) is the nonbounding spin structure.
Equivalently, for the inverse determinant line, the exponent sign is
reversed:
\begin{equation}
\begin{aligned}
  \operatorname{Hol}_{\gamma,\sigma_{\rm b}}(L)
  &=
  \lim_{\varepsilon\to0}
  \exp\!\left(2\pi\ii\,\xi(B_{\gamma,\varepsilon}^{\sigma_{\rm b}})\right),
  \\
  \operatorname{Hol}_{\gamma,\sigma_{\rm nb}}(L)
  &=
  (-1)^{\operatorname{Ind}(D^+)}
  \lim_{\varepsilon\to0}
  \exp\!\left(2\pi\ii\,\xi(B_{\gamma,\varepsilon}^{\sigma_{\rm nb}})\right).
\end{aligned}
  \label{eq:bismut-freed-inverse-holonomy}
\end{equation}
with the same adiabatic family.  The parity factor is the supertrace sign on
the vertical determinant superline.  The limit is invariant under
orientation-preserving reparametrization of \(\gamma\), while the finite
\(\varepsilon\) eta phase need not be.
Its curvature is the fiber integral over \(M\) of the degree
\((\dim M+2)\) component of
\(\widehat A(TM)\operatorname{ch}(E)\).
\end{quotedtheorem}

\paragraph{Local connection forms and holonomy coordinates.}
The theorem supplies the analytic determinant-line connection.  Before using
its particular curvature, we record the elementary line-bundle bookkeeping
that turns such a connection into a global-anomaly phase.  Let
\(\{U_i\}\) be a cover of a patch of \(\mathcal B\), and choose nonzero local
frames \(e_i\) of the determinant line.  Write the normalized connection
one-form in the convention
\[
  \nabla e_i=-2\pi\ii\,a_i\otimes e_i,
  \qquad a_i\in\Omega^1(U_i;\mathbb R).
\]
On an overlap \(U_i\cap U_j\), write
\[
  e_j=\exp(2\pi\ii f_{ij})\,e_i,
  \qquad f_{ij}:U_i\cap U_j\longrightarrow \mathbb R/\mathbb Z .
\]
After choosing a real lift of \(f_{ij}\) on a contractible part of the
overlap, substituting this transition rule into the connection gives
\begin{equation}
  a_j=a_i-df_{ij}.
  \label{eq:det-line-local-connection-transition}
\end{equation}
Therefore \(da_i=da_j\) on overlaps, and the curvature is represented by a
single global closed two-form.  In the Bismut--Freed determinant line this
global two-form is the family-index form stated in
Theorem~\ref{qthm:bismut-freed-holonomy}.

Now let a loop \(\gamma\) be subdivided into arcs
\(\gamma_r:[t_r,t_{r+1}]\to U_{i_r}\), with
\(\gamma(t_N)=\gamma(t_0)\), and let
\(\gamma(t_{r+1})\in U_{i_r}\cap U_{i_{r+1}}\).  Parallel transport along the
arc written in the \(i_r\)-frame contributes
\(\exp(2\pi\ii\int_{\gamma_r}a_{i_r})\).  At the next vertex, the change of
frame from \(e_{i_r}\) to \(e_{i_{r+1}}\) contributes
\(\exp(-2\pi\ii f_{i_r i_{r+1}}(\gamma(t_{r+1})))\).  Hence the holonomy is
\begin{equation}
  \operatorname{Hol}_\gamma(\nabla)
  =
  \exp\left[
    2\pi\ii\left(
      \sum_r\int_{\gamma_r}a_{i_r}
      -
      \sum_r f_{i_r i_{r+1}}(\gamma(t_{r+1}))
    \right)
  \right].
  \label{eq:det-line-cech-holonomy}
\end{equation}
Changing frames by \(e_i\mapsto \exp(2\pi\ii h_i)e_i\) sends
\[
  a_i\mapsto a_i-dh_i,\qquad
  f_{ij}\mapsto f_{ij}+h_j-h_i .
\]
The changes in the arc integrals and the vertex transition terms telescope in
\eqref{eq:det-line-cech-holonomy}.  Thus the closed-loop holonomy is
independent of the local frame.  On a contractible loop contained in a
single patch, \eqref{eq:det-line-cech-holonomy} reduces by Stokes to the
exponential of the curvature integral over a spanning disk.  On a
noncontractible loop the curvature may vanish while
\eqref{eq:det-line-cech-holonomy} remains nontrivial; this is exactly the
flat global-anomaly possibility later detected by based-loop holonomy in the
quotient background groupoid.

\paragraph{Mechanism of determinant-line transport.}
The determinant line over \(\mathcal B\) has local trivializations obtained by
choosing a spectral cut for \(D_b^-D_b^+\) and zeta-regularizing the nonzero
eigenvalue product.  Changing the cut changes the trivialization by the finite
determinant of the eigenmodes crossing the cut, so these local charts glue to
the Quillen determinant line.  The Bismut--Freed connection is the connection
whose local connection form is the finite part of the variation of this
regularized determinant, corrected by the heat-kernel counterterm that makes
the answer independent of the spectral cut.  Its curvature is therefore local:
it is the degree-two form on \(\mathcal B\) obtained by integrating
\([\widehat A(TM)\operatorname{ch}(E)]_{\dim M+2}\) over the physical
manifold \(M\).

The spectral-cut chart can be described without suppressing the finite
transition map.  Fix a positive number \(a\) which is not in the spectrum of
\(K_b^+=D_b^-D_b^+\).  Let
\[
  E_{<a}^+(b)
  =
  \bigoplus_{\lambda<a}\ker(K_b^+-\lambda),
  \qquad
  E_{<a}^-(b)
  =
  D_b^+E_{(0,a)}^+(b)\oplus\ker D_b^- .
\]
Here \(E_{(0,a)}^+\) excludes the zero modes.  The Fredholm determinant line is
represented in this chart by
\[
  L_a(b)
  =
  \det(E_{<a}^+(b))\otimes \det(E_{<a}^-(b))^* .
\]
If \(a<c\) are two cuts avoiding the spectrum, write
\[
  E_{[a,c)}^\pm(b)
  =
  E_{<c}^\pm(b)\ominus E_{<a}^\pm(b).
\]
On this finite-dimensional window, \(D_b^+\) is an isomorphism
\[
  D_{[a,c)}^+(b):
  E_{[a,c)}^+(b)\longrightarrow E_{[a,c)}^-(b),
\]
and the transition from the \(a\)-chart to the \(c\)-chart is multiplication by
the determinant line element of this isomorphism:
\[
  \det D_{[a,c)}^+(b)
  \in
  \det E_{[a,c)}^-(b)\otimes \det E_{[a,c)}^+(b)^* .
\]
For three cuts \(a<c<d\), the window splits as
\[
  E_{[a,d)}^\pm
  =
  E_{[a,c)}^\pm\oplus E_{[c,d)}^\pm,
\]
and determinant multiplicativity gives
\[
  \det D_{[a,d)}^+
  =
  \det D_{[c,d)}^+\,
  \det D_{[a,c)}^+ .
\]
This cocycle identity is the finite-dimensional algebra behind the gluing of
the local determinant-line charts.  The zeta-regularized high-mode determinant
and the heat-kernel counterterm supply the infinite-dimensional continuation of
the same transition rule.

\paragraph{Compatibility with eta jumps.}
The determinant-line chart transition and the reduced eta invariant describe
the same zero-mode bookkeeping from two different sides.  Consider first one
real eigenvalue \(\lambda(t)\) of a self-adjoint tangential operator \(B_t\)
crossing through zero from negative to positive, with all other eigenvalues
kept fixed away from zero.  Its contribution to the reduced eta invariant is
\[
  \xi_\lambda(t)
  =
  \begin{cases}
    -\frac12, & \lambda(t)<0,\\[2mm]
    \frac12, & \lambda(t)=0,\\[2mm]
    \frac12, & \lambda(t)>0,
  \end{cases}
\]
where the middle value comes from the kernel term \(h(B_t)\).  The
one-sided change of \(\xi\) is therefore \(+1\), and a downward crossing gives
\(-1\).  Thus the exponentiated phase
\(\exp(-2\pi\ii\,\xi)\) is continuous across such an integer jump.

In the determinant-line construction the same event appears as a change of
low-mode chart.  Near a kernel jump, the zeta determinant of the complement
and the zero-mode factor carried by the finite-dimensional determinant line
\(L_a(b)\) together form the fermion determinant vector.  When the cut is
moved across a small eigenvalue window, the local scalar determinant changes
by the finite determinant \(\det D^+_{[a,c)}\), while the basis of the
determinant-line chart changes by the inverse transition element.  Their
product is a vector in the Quillen line and is independent of the auxiliary
spectral cut.  The integer spectral-flow term in the eta description is
precisely the phase shadow of this finite chart transition.  Consequently the
smooth local curvature, the finite spectral-flow jumps, and the
mapping-torus eta phase are the curvature, transition, and holonomy data of
the same determinant line.

For a loop \(\gamma\), the holonomy problem is converted into an adiabatic
index problem on the odd-dimensional mapping torus \(Y_\gamma\).  The local
curvature contribution becomes the transgression of the family index density,
while the nonlocal part left by the adiabatic limit is the reduced eta
invariant in \eqref{eq:bismut-freed-holonomy}.  A finite-metric eta phase on
the same mapping torus is an auxiliary Riemannian number; it can change when
the loop is reparametrized and is not itself determinant-line transport.
Thus the local anomaly polynomial is the curvature of the anomaly line, and
a global anomaly is the adiabatic holonomy of the same line around a closed
path in the background groupoid.  Cancellation of the local anomaly means
that the total anomaly line is flat; cancellation of the global anomaly
requires the holonomy of that flat line to be \(1\) on every closed loop.

\paragraph{Finite gauge transformation as determinant-line transport.}
Let \(L=\operatorname{Det}D\) for the moment.  It carries two distinct pieces
of structure.  First, covariance of the Dirac family under a gauge
transformation \(g\) gives a canonical equivariant line map
\[
  \rho_g(A):L_A\longrightarrow L_{A^g},
  \qquad
  \rho_h(A^g)\rho_g(A)=\rho_{gh}(A).
\]
This map is functorial in the gauge action and is not chosen by selecting a
path in background-field space.  Second, the Bismut--Freed connection gives
parallel transport
\[
  P_\gamma:L_A\longrightarrow L_{A^g}
\]
along any path \(\gamma:[0,1]\to\mathcal B\) with
\(\gamma(0)=A\) and \(\gamma(1)=A^g\).

The closed loop in the quotient background stack is the pair
\((\gamma,g)\), not the gauge arrow \(g\) alone.  With the orientation
convention used here, its determinant-line phase is the endomorphism
\begin{equation}
  \Theta_A(\gamma,g)
  =
  \rho_g(A)^{-1}P_\gamma
  \in\operatorname{Aut}(L_A)\cong\mathbb C^\times .
  \label{eq:quotient-loop-phase-rho-transport}
\end{equation}
Geometrically the same quotient loop is represented by the mapping torus
\begin{equation}
  Y_g
  =
  ([0,1]\times M)/\bigl((1,x)\sim(0,gx)\bigr).
  \label{eq:gauge-mapping-torus}
\end{equation}
The Bismut--Freed holonomy formula applied to this adiabatic mapping torus
computes the scalar in \eqref{eq:quotient-loop-phase-rho-transport}, with the
chosen circle spin structure and determinant-line variance fixed above.  A
phase different from \(1\) means that Bismut--Freed transport and the
canonical gauge identification disagree; hence the determinant vector is not
a single-valued function on the quotient background stack.

The separation between \(\rho\) and \(P_\gamma\) also gives the composition
law.  If \(\delta\) is a path from \(A^g\) to \(A^{gh}\), then
\(P_{\gamma*\delta}=P_\delta P_\gamma\) and
\(\rho_{gh}(A)=\rho_h(A^g)\rho_g(A)\).  Therefore, as scalars on the
one-dimensional fibers,
\begin{equation}
  \Theta_A(\gamma*\delta,gh)
  =
  \Theta_A(\gamma,g)\,
  \Theta_{A^g}(\delta,h).
  \label{eq:quotient-loop-phase-composition}
\end{equation}
If \(\gamma_0\) and \(\gamma_1\) are two paths from \(A\) to \(A^g\), then
\(\rho_g(A)\) is unchanged, while
\[
  \Theta_A(\gamma_1,g)\Theta_A(\gamma_0,g)^{-1}
\]
is the Bismut--Freed holonomy around the closed path obtained by concatenating
\(\gamma_0\) and \(\gamma_1\) with one orientation reversed, with orientation
fixed by the displayed ratio.  For a contractible change swept out by a
surface \(\Sigma\subset\mathcal B\), this ratio is controlled by the curvature
integral
\(\exp(2\pi\ii\int_\Sigma\Omega_{\mathrm{BF}})\).  After local anomaly
cancellation the remaining line is flat, so the quotient-loop phase depends
only on the corresponding homotopy class in the quotient stack.

\paragraph{Descent to the quotient background groupoid.}
This last statement is the precise descent criterion for a global gauge
anomaly.  Let \(\mathcal B\) be the space of smooth background fields before
quotienting, and let \(\mathcal G\) be the group of gauge transformations
acting on it.  The quotient is the action groupoid
\[
  \mathcal B//\mathcal G ,
\]
whose objects are \(A\in\mathcal B\) and whose morphisms
\[
  A\longrightarrow A^g
\]
are labelled by \(g\in\mathcal G\).  A fermion determinant line over this
action groupoid consists of a line \(L_A\) over every object together with
the canonical gauge-equivariant isomorphisms
\[
  \rho_g(A):L_A\longrightarrow L_{A^g}
\]
compatible with composition.  The Bismut--Freed connection is extra
geometric structure on this equivariant line; its path transports are used to
evaluate quotient-stack loops through
\eqref{eq:quotient-loop-phase-rho-transport}, but they are not the strict
action-groupoid morphisms.

A partition function is a function on the quotient only after choosing
nonzero vectors \(s_A\in L_A\) satisfying
\begin{equation}
  \rho_g(A)s_A=s_{A^g}
  \qquad\hbox{for every }(A,g).
  \label{eq:anomaly-line-equivariant-trivialization}
\end{equation}
If one fixes a local trivialization of \(L\), the same condition is expressed
by a \(U(1)\)-valued groupoid \(1\)-cocycle
\[
  \alpha(A,g)\in U(1),
  \qquad
  \rho_g(A)e_A=\alpha(A,g)e_{A^g},
\]
obeying
\begin{equation}
  \alpha(A,gh)=\alpha(A,g)\,\alpha(A^g,h).
  \label{eq:anomaly-line-groupoid-cocycle}
\end{equation}
A change of local trivialization \(e_A\mapsto\beta(A)e_A\) changes
\(\alpha\) by the coboundary
\begin{equation}
  \alpha(A,g)\longmapsto
  \frac{\beta(A)}{\beta(A^g)}\,\alpha(A,g).
  \label{eq:anomaly-line-groupoid-coboundary}
\end{equation}
Thus the obstruction is the cohomology class of this groupoid cocycle.  On a
stabilizer arrow \(g\in\mathcal G_A\), the value \(\alpha(A,g)\) is
invariant under \eqref{eq:anomaly-line-groupoid-coboundary}.  In the
orientation convention of \eqref{eq:quotient-loop-phase-rho-transport}, the
constant-path quotient-loop phase is its inverse.  Such stabilizer phases are
important, but they are only special loops in the homotopy quotient: a
general loop is a path \(\gamma:A\to A^g\) together with the endpoint gauge
identification \(g\), and its phase also includes Bismut--Freed parallel
transport along \(\gamma\).  Tensoring independent fermion systems tensors
their anomaly lines, so the cocycles multiply and the logarithmic anomaly
exponents add.

This based-loop statement has a useful fixed-background form.  Let
\[
  \mathcal G_A=\{g\in\mathcal G:A^g=A\}
\]
be the stabilizer of a background.  Restricting
\eqref{eq:anomaly-line-groupoid-cocycle} to \(g,h\in\mathcal G_A\) gives a
character
\[
  \chi_A(g)=\alpha(A,g),
  \qquad
  \chi_A(gh)=\chi_A(g)\chi_A(h).
\]
The coboundary factor in
\eqref{eq:anomaly-line-groupoid-coboundary} is trivial on the stabilizer,
because \(A^g=A\), so \(\chi_A\) is not changed by a local choice of frame.
If an equivariant nonzero section \(s_A\) satisfying
\eqref{eq:anomaly-line-equivariant-trivialization} exists, then
\(\chi_A(g)=1\) for every stabilizer element.  The converse is a concrete
descent theorem in the finite flat action-groupoid model.

The following proposition has this restricted scope: it is a finite
algebraic model for a flat discrete reduction of the anomaly line, after the
local Bismut--Freed curvature obstruction has been cancelled or otherwise
removed.  It does not encode nonflat path holonomies before that reduction.

\begin{proposition}[Finite flat action-groupoid descent criterion]
\label{prop:finite-action-groupoid-descent-criterion}
Let a finite group \(G\) act on a finite set \(\mathcal B_0\) of background
representatives on the right, and write the action groupoid as
\(\mathcal B_0//G\).  Let
\[
  a(A,g)\in\mathbb R/\mathbb Z
\]
be an additive representative of a flat line cocycle,
\begin{equation}
  a(A,gh)=a(A,g)+a(A^g,h).
  \label{eq:finite-action-groupoid-additive-cocycle}
\end{equation}
Then \(a\) is a coboundary,
\begin{equation}
  a(A,g)=b(A^g)-b(A)
  \label{eq:finite-action-groupoid-coboundary}
\end{equation}
for some \(0\)-cochain \(b:\mathcal B_0\to\mathbb R/\mathbb Z\), if and only
if its stabilizer character vanishes on every orbit,
\begin{equation}
  a(A,h)=0
  \qquad
  \text{for all }h\in G_A .
  \label{eq:finite-action-groupoid-stabilizer-vanishing}
\end{equation}
Equivalently, the multiplicative line cocycle
\(\alpha(A,g)=\exp(2\pi\ii a(A,g))\) is trivializable by a choice of local
frames exactly when all finite action-groupoid based-loop holonomies are
\(1\).
\end{proposition}

\begin{proof}
If \(a(A,g)=b(A^g)-b(A)\), then for a stabilizer element \(h\in G_A\) one has
\(A^h=A\), so \(a(A,h)=b(A)-b(A)=0\).  Conversely, choose one representative
\(A_o\) in each orbit and set \(b(A_o)=0\).  If \(A=A_o^g\), define
\[
  b(A)=a(A_o,g).
\]
This is independent of the chosen \(g\).  Indeed, if \(A_o^g=A_o^{g'}\), then
\(g'=hg\) for some \(h\in G_{A_o}\), and the cocycle identity gives
\[
  a(A_o,g')=a(A_o,hg)=a(A_o,h)+a(A_o^h,g)=a(A_o,g),
\]
because \(A_o^h=A_o\) and the stabilizer character vanishes.  For an arbitrary
arrow \(A=A_o^g\to A^k=A_o^{gk}\), the same cocycle identity gives
\[
  b(A^k)-b(A)
  =
  a(A_o,gk)-a(A_o,g)
  =
  a(A_o^g,k)
  =
  a(A,k).
\]
Thus the required \(b\) exists.  Exponentiating \(a\) and \(b\) gives the
multiplicative statement and matches the frame-change law
\eqref{eq:anomaly-line-groupoid-coboundary}.
\end{proof}

This finite theorem is the algebraic core of the flat descent problem, not a
substitute for the smooth connection.  For connected transformations the
Bismut--Freed curvature and the local descent calculation control how
\(\Theta_A(\gamma,g)\) changes under contractible changes of \(\gamma\).
Only after that curvature has been cancelled does the remaining flat line
reduce to a discrete cocycle of the kind in the proposition.  A nontrivial
stabilizer character at one background is already an obstruction, but it does
not replace the need to test general quotient-stack loops when the homotopy
quotient has closed paths not represented by stabilizer arrows.

\paragraph{Dual anomaly lines and the cancellation criterion.}
Let \(L\) be an anomaly line over \(\mathcal B//\mathcal G\).  Its dual line
\(L^\vee\) has fibers
\[
  L_A^\vee=\operatorname{Hom}(L_A,\mathbb C).
\]
If \(\rho_g(A):L_A\to L_{A^g}\) is the equivariant map of \(L\), then the
equivariant map of \(L^\vee\) is the inverse dual map
\[
  \rho_g^\vee(A)
  =
  \bigl(\rho_g(A)^{-1}\bigr)^{*}
  :
  L_A^\vee\longrightarrow L_{A^g}^\vee .
\]
In a local trivialization, if \(L\) is represented by the cocycle
\(\alpha(A,g)\), then \(L^\vee\) is represented by
\(\alpha(A,g)^{-1}\).  Hence
\begin{equation}
  \alpha_{L\otimes L^\vee}(A,g)
  =
  \alpha(A,g)\alpha(A,g)^{-1}
  =
  1 .
  \label{eq:anomaly-line-dual-cancellation-cocycle}
\end{equation}
The tensor product \(L\otimes L^\vee\) is therefore canonically trivialized by
the evaluation pairing.  More generally, anomaly cancellation for a collection
of fermion systems is the existence of a groupoid-equivariant trivialization
of the tensor product of their determinant or Pfaffian lines.  Vanishing of
the local curvature only says that this tensor product is flat.  It still
must have trivial based-loop holonomy, because a nontrivial flat line gives a
well-defined local anomaly density equal to zero while failing
\eqref{eq:anomaly-line-equivariant-trivialization} on a noncontractible loop
of backgrounds.  For real Pfaffian lines the inverse line is again a real
line; the statement becomes multiplication of signs.  Thus two identical
four-dimensional \(SU(2)\) Weyl doublets cancel the Witten sign because
\((-1)^2=1\), whereas one doublet gives a nontrivial flat Pfaffian line over
the large-gauge loop.

For an infinitesimal gauge transformation connected to the identity, the
linearization of \(\alpha\) is the consistent local anomaly derived in
Chapter~\ref{ch:volume-ii-20}.  For a large gauge transformation such as the
generator of \(\pi_4(SU(2))\), the local curvature may vanish while the based
loop value of \(\alpha\) is nontrivial.  This is the logical distinction
between local anomaly cancellation and global anomaly cancellation.

\section{Pfaffian Lines and Mod-Two Indices}

Some fermion systems have a real or pseudoreal structure that replaces the
complex determinant line by a Pfaffian line.  The four-dimensional
\(SU(2)\) fundamental Weyl fermion is the basic example.  The fundamental
representation is pseudoreal, and the spin representation supplies an
anti-linear symmetry that makes the Euclidean quadratic form skew on a real
vector space.  The path integral is then a Pfaffian.  Its possible global
anomaly is a sign.
The Pfaffian line is also a superline, but its parity is a separate real
Fredholm parity, not the complex chiral index in
\eqref{eq:det-superline-parity}.  For the real skew Fredholm problem used
below we write
\[
  p_{\rm Pf}(\operatorname{Pf}D)
  =
  \operatorname{Ind}_2(D_{\mathbb R})
  =
  \dim_{\mathbb R}\ker D_{\mathbb R}
  \pmod 2 .
\]
This is the parity used for real Pfaffian gluing and for the Witten
mapping-torus sign.

\begin{definition}[Mod-two index]
\label{def:mod-two-index}
Let \(Y\) be a closed spin manifold of dimension \(8k+5\), and let \(E\) be a
pseudoreal bundle for which the coupled Dirac operator has the corresponding
real skew structure.  The mod-two index is
\begin{equation}
  \operatorname{Ind}_2(B_Y)
  =
  \dim\ker B_Y
  \pmod 2 .
  \label{eq:mod-two-index-definition}
\end{equation}
The parity is stable under continuous deformations preserving the real
structure.  The reason is that eigenvalues away from zero occur in pairs
\(\lambda,-\lambda\) for the real
skew structure relevant to the Pfaffian.  Under a deformation, nonzero
eigenvalues can enter or leave the kernel only through such paired crossings.
Thus the dimension of the kernel can change only by an even integer, and
\(\dim\ker B_Y\) modulo \(2\) is invariant.
\end{definition}

For such a Pfaffian line the Dai--Freed holonomy specializes to
\begin{equation}
  \operatorname{Hol}_\gamma(\operatorname{Pf}D)
  =
  (-1)^{\operatorname{Ind}_2(B_{Y_\gamma})}.
  \label{eq:pfaffian-mod-two-holonomy}
\end{equation}
This formula is the sign analogue of
\eqref{eq:bismut-freed-holonomy}.

\paragraph{From determinant phases to Pfaffian signs.}
The Pfaffian line is a real line, and its holonomy has values in
\(\{\pm1\}\).  In a real skew Fredholm family, the nonzero spectrum of the
self-adjoint representative obtained by multiplying the skew operator by
\(\ii\) is symmetric about zero.  Along a closed path, the orientation of the
Pfaffian line changes precisely when an odd number of eigenvalue pairs passes
through zero.  The parity of this crossing number is the mod-two index of the
Dirac operator on the mapping torus.  Formula
\eqref{eq:pfaffian-mod-two-holonomy} is therefore the real-line analogue of
the Bismut--Freed eta phase: the exponent is no longer a reduced eta invariant
modulo integers, but a \(\mathbb Z_2\)-valued index.

\paragraph{Finite skew model for the Pfaffian line.}
The sign mechanism can be seen without functional analysis in a finite
oriented real vector space.  Let \(V=\mathbb R^2\) with oriented basis
\((e_1,e_2)\), and let
\[
  q_a=a\,e^1\wedge e^2
\]
be a skew bilinear form.  Its Pfaffian in this oriented basis is
\begin{equation}
  \operatorname{pf}(q_a)=a,
  \qquad
  \det(q_a)=a^2 .
  \label{eq:finite-pfaffian-block}
\end{equation}
Thus the determinant sees only \(|a|\), while the Pfaffian remembers the
orientation of the real line generated by \(q_a\).  If \(a(t)\) crosses
through zero once, the real Pfaffian coordinate changes sign.  At the
crossing the skew form has a two-dimensional kernel; away from the crossing
the associated self-adjoint block has a paired nonzero spectrum.  This is the
finite-dimensional version of the statement that a real skew Fredholm family
changes Pfaffian orientation only through paired zero-mode crossings.

For a direct sum of such blocks,
\[
  (V,q)=\bigoplus_{r=1}^{N}(\mathbb R^2,q_{a_r}),
\]
with the product orientation, the Pfaffian is multiplicative:
\begin{equation}
  \operatorname{pf}(q)=\prod_{r=1}^{N}a_r .
  \label{eq:finite-pfaffian-block-product}
\end{equation}
Consequently the holonomy of the real Pfaffian line around a closed family is
the parity of the number of block signs that change:
\begin{equation}
  \operatorname{Hol}(\operatorname{Pf})
  =
  (-1)^{N_{\rm cross}\bmod2}.
  \label{eq:finite-pfaffian-crossing-parity}
\end{equation}
The mod-two index in \eqref{eq:pfaffian-mod-two-holonomy} is the
infinite-dimensional, elliptic version of this finite orientation parity.
The mapping-torus Dirac operator supplies a canonical way to count the parity
of the zero-mode crossings without choosing coordinates along the loop of
backgrounds.

\section{\texorpdfstring{The Witten \(SU(2)\) Global Anomaly}{The Witten SU(2) Global Anomaly}}

Let \(M=S^4\), and let \(g:S^4\to SU(2)\) represent the nontrivial element of
\[
  \pi_4(SU(2))\simeq\mathbb Z_2 .
\]
Form the mapping torus \(Y_g\) as in \eqref{eq:gauge-mapping-torus}.  A
single left-handed Weyl fermion in the fundamental representation of
\(SU(2)\) has Pfaffian holonomy
\begin{equation}
  \operatorname{Hol}_g(\operatorname{Pf}D)=-1 .
  \label{eq:witten-su2-fundamental-sign}
\end{equation}
Equivalently,
\begin{equation}
  \operatorname{Ind}_2(B_{Y_g,\square})=1 .
  \label{eq:witten-su2-fundamental-mod-two-index}
\end{equation}

\begin{quotedtheorem}[Mod-two index theorem for the Witten mapping torus]
\label{qthm:witten-su2-mod-two-index}
\emph{Source trace.}  The global \(SU(2)\) anomaly and its mod-two
mapping-torus index formulation trace to Witten's anomaly computation, with
the determinant/Pfaffian-line language aligned with the Dai--Freed framework
\cite{Witten1982,DaiFreed1994}.
Let \(R_j\) be the irreducible \(SU(2)\) representation of isospin
\(j\in\frac12\mathbb Z_{\geq0}\), and set \(n=2j\).  With the trace-delta
normalization of Chapter~\ref{ch:curved-background-gauge-fields-index-theory},
\begin{equation}
  \operatorname{tr}_{R_j}(t^a t^b)
  =
  T_\Delta(R_j)\delta^{ab},
  \qquad
  T_\Delta(R_j)
  =
  \frac{n(n+1)(n+2)}6 .
  \label{eq:su2-trace-delta-dynkin-index}
\end{equation}
For the mapping torus of the generator of \(\pi_4(SU(2))\),
\begin{equation}
  \operatorname{Ind}_2(B_{Y_g,R_j})
  \equiv
  T_\Delta(R_j)
  \pmod2 .
  \label{eq:witten-su2-index-parity-dynkin}
\end{equation}
\end{quotedtheorem}

\paragraph{Clutching datum behind the mapping torus.}
The QFT object to which Theorem~\ref{qthm:witten-su2-mod-two-index} is
applied is the coupled Dirac operator on the associated bundle over the
mapping torus determined by \(g:S^4\to SU(2)\).  Start with the
trivial principal \(SU(2)\)-bundle over \(S^4\).  In a trivialization write a
point as \((x,u)\), \(x\in S^4\), \(u\in SU(2)\).  The clutching bundle over
\(Y_g=S^1\times_g S^4\) is
\begin{equation}
  P_g
  =
  ([0,1]\times S^4\times SU(2))/
  \bigl((1,x,u)\sim(0,x,g(x)u)\bigr).
  \label{eq:witten-su2-clutching-bundle}
\end{equation}
For a representation \(R\), the bundle entering the Dirac operator is
\[
  E_{g,R}=P_g\times_{SU(2)}R .
\]
Choose a smooth path of connections \(A_t\) from a connection \(A\) to its
gauge transform \(A^g\), product near \(t=0,1\).  The connection \(A_t\) on
the cylinder descends through the clutching relation to a connection on
\(P_g\).  Changing the interpolation \(A_t\) changes the mapping-torus
connection through a homotopy of connections and hence does not change the
mod-two index.

Two pieces of the setup precede the global mod-two index formula.  First, if
\(g_s\), \(0\leq s\leq1\), is a homotopy of gauge
transformations, the same clutching construction over
\([0,1]_s\times[0,1]_t\times S^4\), with the \(t\)-ends identified by
\(g_s\), gives a six-dimensional bordism between
\((Y_{g_0},P_{g_0})\) and \((Y_{g_1},P_{g_1})\).  The mod-two index is a
bordism invariant, so the sign depends only on the homotopy class
\([g]\in\pi_4(SU(2))\).  Second, concatenating two gauge transformations
amounts to gluing the corresponding mapping cylinders along an intermediate
copy of \(S^4\); additivity of the real mod-two index under gluing gives
\begin{equation}
  \operatorname{Ind}_2(B_{Y_{g_0g_1},R})
  =
  \operatorname{Ind}_2(B_{Y_{g_0},R})
  +
  \operatorname{Ind}_2(B_{Y_{g_1},R})
  \pmod2 .
  \label{eq:witten-su2-mapping-torus-additivity}
\end{equation}
Thus the Pfaffian sign is a character of the component group of the gauge
group.  For \(SU(2)\) on \(S^4\), this component group is
\(\pi_4(SU(2))\simeq\mathbb Z_2\).  The nontrivial content of
Theorem~\ref{qthm:witten-su2-mod-two-index} is the value of this character on
the generator, and its representation dependence.  The clutching construction
above is the part of the global-anomaly argument that belongs directly to the
QFT background-field setup.

\paragraph{Trace-delta normalization in the \(SU(2)\) representation ring.}
The arithmetic in Theorem~\ref{qthm:witten-su2-mod-two-index} uses the same
trace convention as the instanton index formula
\eqref{eq:index-trace-delta-instanton-number}.  Let \(J_a\) be the usual
Hermitian spin-\(j\) generators with
\([J_a,J_b]=\ii\epsilon^c{}_{ab}J_c\) and \(J_3\)-weights
\[
  m=-j,-j+1,\ldots,j .
\]
Rotational invariance of the irreducible representation gives
\[
  \operatorname{tr}_{R_j}(J_aJ_b)=C_j\delta_{ab}.
\]
Taking \(a=b=3\) fixes
\begin{equation}
  C_j
  =
  \sum_{m=-j}^{j}m^2
  =
  \frac13j(j+1)(2j+1).
  \label{eq:su2-spin-j-half-trace-index}
\end{equation}
The monograph's trace-delta basis is \(t_a=\sqrt2\,J_a\), since the
fundamental spin-\(\frac12\) representation then has
\(\operatorname{tr}_{\square}(t_at_b)=\delta_{ab}\).  Therefore
\begin{equation}
  T_\Delta(R_j)
  =
  2C_j
  =
  \frac23j(j+1)(2j+1)
  =
  \frac{n(n+1)(n+2)}6,
  \qquad n=2j .
  \label{eq:su2-spin-j-trace-delta-derived}
\end{equation}
This derivation fixes the normalization of the integer that appears in the
mod-two index theorem; it is not an independent anomaly calculation.

The parity of \eqref{eq:su2-spin-j-trace-delta-derived} is equally explicit:
\[
  T_\Delta(R_j)=\binom{n+2}{3}.
\]
Lucas' binary criterion for binomial coefficients modulo \(2\) says that
\(\binom N3\) is odd exactly when the two lowest binary digits of \(N\) are
both \(1\).  With \(N=n+2\), this is
\begin{equation}
  n+2\equiv3\pmod4,
  \qquad\hbox{equivalently}\qquad
  n=2j\equiv1\pmod4 .
  \label{eq:su2-witten-parity-arithmetic}
\end{equation}
Thus the quoted global-analysis input is the mod-two-index equality
\eqref{eq:witten-su2-index-parity-dynkin}; the representation normalization
and the \(2j\equiv1\pmod4\) criterion are ordinary finite-dimensional
\(SU(2)\) arithmetic in the chapter's trace convention.

\paragraph{Topological content of the mapping-torus index.}
The theorem is the point where topology enters the four-dimensional anomaly.
The gauge transformation \(g\) representing the nontrivial element of
\(\pi_4(SU(2))\) produces a spin five-manifold \(Y_g\) equipped with an
\(SU(2)\) bundle, and the pair \((Y_g,E_{R_j})\) defines an element tested by
the mod-two Dirac index.  For the fundamental representation this element has
odd mod-two index; this is the genuinely global datum invisible to the local
four-dimensional cubic anomaly polynomial.  The theorem further identifies the
parity for a general irreducible representation with the trace-delta Dynkin
index modulo \(2\), giving \eqref{eq:witten-su2-index-parity-dynkin}.  Its
proof uses the naturality of the coupled Dirac operator under representation
operations together with a five-dimensional real-index computation for the
generator of the \(\pi_4(SU(2))\) mapping torus.  The QFT analysis below uses
the result only through the explicit parity formula displayed in the theorem.

Once \eqref{eq:witten-su2-index-parity-dynkin} is fixed, the remaining anomaly
criterion is ordinary arithmetic in the same trace convention used for the
local anomaly polynomial.

The isospin-\(j\) Weyl fermion has the original spin \(SU(2)\) global anomaly
precisely when \(2j\equiv1\pmod4\).  A collection of Weyl fermions in
representations \(R_{j_\alpha}\) is free of this anomaly precisely when
\begin{equation}
  \sum_\alpha T_\Delta(R_{j_\alpha})
  \equiv0
  \pmod2 .
  \label{eq:su2-global-anomaly-cancellation-condition}
\end{equation}
Write \(n=2j\).  The integer
\[
  T_\Delta(R_j)=\frac{n(n+1)(n+2)}6
\]
is the binomial coefficient \(\binom{n+2}{3}\).  Its parity is given by
Lucas' theorem: \(\binom{N}{3}\) is odd exactly when the binary digits of
\(3\) are contained in those of \(N\).  Since \(3\) has binary expansion
\(\cdots 0011\), this means \(N=n+2\equiv3\pmod4\), or
\[
  n\equiv1\pmod4 .
\]
Equation~\eqref{eq:witten-su2-index-parity-dynkin} then gives the Pfaffian
sign.  For several Weyl fermions the Pfaffian line is the tensor product of
the individual Pfaffian lines, so the signs multiply and the exponents add
modulo \(2\).

\begin{example}[First few \(SU(2)\) representations]
\label{ex:su2-global-anomaly-first-representations}
The sequence of trace-delta indices is
\[
\begin{array}{c|ccccc}
j & \frac12 & 1 & \frac32 & 2 & \frac52\\
\hline
T_\Delta(R_j) & 1 & 4 & 10 & 20 & 35 .
\end{array}
\]
Thus a single doublet has the anomaly, a single triplet does not, a
single four-dimensional representation does not, and a single
six-dimensional representation has the anomaly.  The vanishing of the
ordinary four-dimensional cubic \(SU(2)\) anomaly does not decide this
question because \(SU(2)\) has no independent symmetric cubic invariant.
\end{example}

\section{Dai--Freed Inflow}

The determinant or Pfaffian of a \(d\)-dimensional fermion is a vector in an
anomaly line over the background groupoid.  A \((d+1)\)-dimensional
invertible field theory trivializes this line when its boundary variation is
the inverse of the fermion-line variation.  In the free-fermion model the
closed \((d+1)\)-manifold partition function is the eta phase
\cite{DaiFreed1994,Freed1995DeterminantRevisited}:
\begin{equation}
  Z_{\mathrm{DF}}(Y)
  =
  \exp\!\left(2\pi\ii\,\xi(B_Y)\right)
  \label{eq:dai-freed-closed-phase}
\end{equation}
for a complex chiral determinant, with the corresponding mod-two sign in the
real Pfaffian case.

If \(Y\) has boundary \(M\), the APS construction assigns not a number but a
vector in the inverse determinant line over the boundary background on \(M\).
Pairing this vector with the boundary fermion determinant gives a number.
Changing the filling changes the trivialization by the closed-manifold phase
obtained by gluing the two fillings.  This is the precise geometric meaning
of anomaly inflow in the eta-invariant model.

\paragraph{Closed phases from bounding manifolds.}
The closed phase \eqref{eq:dai-freed-closed-phase} is constrained by the APS
formula whenever the odd-dimensional pair \((Y,E)\) bounds an even-dimensional
pair \((X,\widetilde E)\) with product structure near \(Y=\partial X\).  In
the convention of \eqref{eq:aps-index-formula},
\[
  \operatorname{Ind}(D_X^{+,\mathrm{APS}})
  =
  \int_X
  \bigl[\widehat A(TX)\operatorname{ch}(\widetilde E)\bigr]_{\dim X}
  -
  \xi(B_Y).
\]
The index is an integer, hence
\begin{equation}
  \exp\!\left(2\pi\ii\,\xi(B_Y)\right)
  =
  \exp\!\left(
    2\pi\ii
    \int_X
    \bigl[\widehat A(TX)\operatorname{ch}(\widetilde E)\bigr]_{\dim X}
  \right)
  \label{eq:dai-freed-closed-phase-bounding-integral}
\end{equation}
for bounding data.  If \(X\) and \(X'\) are two such fillings, gluing
\(X\) to \(-X'\) gives a closed even-dimensional spin manifold.  The closed
Atiyah--Singer theorem says that the difference of the two integrals in
\eqref{eq:dai-freed-closed-phase-bounding-integral} is an integer.  Thus the
phase is independent of the chosen filling, even though neither integral is
itself a canonical real number modulo ordinary equality.  This is the point
at which the bulk local index density and the boundary eta invariant are tied
to the same \(U(1)\)-valued object.

\paragraph{Boundary vector and gluing law.}
Let \(Y\) now be an odd-dimensional spin manifold with boundary \(M\), and
let all geometric data be product type near \(M\).  The Dai--Freed
construction assigns a vector
\[
  Z_{\mathrm{DF}}(Y)
  \in
  L_M^\vee,
\]
where \(L_M\) is the determinant or Pfaffian anomaly superline of the
boundary fermion system.  For a complex chiral determinant
\(p(M)=\operatorname{Ind}(D_M^+)\bmod2\), and for a real Pfaffian line
\(p(M)=p_{\rm Pf}(M)\).  Orientation reversal is dualization in the graded
category,
\begin{equation}
  L_{-M}\simeq L_M^\vee,
  \qquad
  p(-M)=p(M).
  \label{eq:graded-orientation-reversal-dual}
\end{equation}
This vector is a trivialization of the boundary anomaly superline, not a
numerical partition function.  The numerical partition function appears only
after pairing it with the boundary fermion determinant vector in \(L_M\).

The interface contraction is the graded trace on the common boundary line.
For homogeneous local dual frames \(\ell^\vee\in L_M^\vee\) and
\(\ell\in L_M\), our convention is
\begin{equation}
  \operatorname{str}_{L_M}(\ell^\vee\otimes\ell)
  =
  (-1)^{p(M)}\ell^\vee(\ell).
  \label{eq:dai-freed-supertrace-interface}
\end{equation}
Thus an odd-index determinant or Pfaffian superline differs from ordinary
evaluation by a minus sign, while an even-index interface reduces to ordinary
line algebra.  For the complex determinant case this is the explicit factor
\((-1)^{\operatorname{Ind}(D_M^+)}\).

If \(Y_0\) and \(Y_1\) have the same boundary \(M\), with the same boundary
background, the ratio between the two boundary trivializations is the closed
phase of the glued odd manifold
\[
  \overline Y
  =
  Y_0\cup_M(-Y_1).
\]
Since \(Z_{\rm DF}(Y_0)\in L_M^\vee\) and
\(Z_{\rm DF}(-Y_1)\in L_M\) by
\eqref{eq:graded-orientation-reversal-dual}, the determinant-line pairing is
the graded gluing identity
\begin{equation}
  \operatorname{str}_{L_M}
  \bigl(Z_{\mathrm{DF}}(Y_0)\otimes Z_{\mathrm{DF}}(-Y_1)\bigr)
  =
  Z_{\mathrm{DF}}\bigl(Y_0\cup_M(-Y_1)\bigr),
  \label{eq:dai-freed-boundary-gluing-law}
\end{equation}
with the same formula for the real Pfaffian superline after replacing
\(p(M)\) by \(p_{\rm Pf}(M)\).  In an ordinary local coordinate the left side
is \((-1)^{p(M)}\) times the ungraded evaluation.  The right side is a number
because the boundary anomaly superlines cancel under gluing.  Associativity
of gluing is therefore not an extra physical postulate: it is the statement
that tensoring one-dimensional superlines in a fixed order and composing
their graded traces gives the same number as gluing all odd-dimensional
pieces first.  This is the finite superline algebra that the eta invariant
realizes analytically.

\begin{proposition}[Dai--Freed boundary pairing cancels the anomaly cocycle]
\label{prop:dai-freed-boundary-pairing-cancels-cocycle}
Let \(A\) be a structured background on \(M\), let \(g\) be a gauge or
background automorphism, and let
\(\alpha(A,g):L_A\to L_{A^g}\) denote the determinant- or Pfaffian-line
transport.  Suppose the boundary fermion functional is a line-valued vector
\(\Psi(A)\in L_A\) with covariance
\begin{equation}
  \Psi(A^g)=\alpha(A,g)\Psi(A),
  \label{eq:boundary-fermion-line-covariance}
\end{equation}
and a Dai--Freed filling \(Y\) supplies a vector
\(Z_{\rm DF}(Y;A)\in L_A^\vee\) with inverse covariance
\begin{equation}
  Z_{\rm DF}(Y;A^g)
  =
  \alpha(A,g)^{-1}Z_{\rm DF}(Y;A).
  \label{eq:dai-freed-inverse-line-covariance}
\end{equation}
Then the paired amplitude
\begin{equation}
  \mathcal A_Y(A)
  =
  \operatorname{str}_{L_A}
  \bigl(Z_{\rm DF}(Y;A)\otimes\Psi(A)\bigr)
  \label{eq:dai-freed-paired-boundary-amplitude}
\end{equation}
is invariant under \(g\).  If \(Y_0\) and \(Y_1\) are two fillings with the
same boundary background, then
\begin{equation}
  \frac{\mathcal A_{Y_0}(A)}{\mathcal A_{Y_1}(A)}
  =
  Z_{\rm DF}\bigl(Y_0\cup_M(-Y_1)\bigr),
  \label{eq:dai-freed-filling-change-closed-phase}
\end{equation}
whenever the denominator is nonzero.  Thus inflow does not make the anomalous
boundary determinant into a scalar by itself; it supplies the inverse line
vector needed to form a scalar, and changing the inverse-line trivialization
is measured by a closed invertible-theory phase.
\end{proposition}

\begin{proof}
Let \(p(A)\) be the parity of the superline \(L_A\), which is preserved by
the line transport.  Work in a local frame \(e_A\) of \(L_A\).  Write
\(\Psi(A)=\psi_A e_A\), \(Z_{\rm DF}(Y;A)=z_A e_A^\vee\), and let the line
transport multiply the frame coordinate by a phase \(\alpha(A,g)\).  The
covariance equations give
\(\psi_{A^g}=\alpha(A,g)\psi_A\) and
\(z_{A^g}=\alpha(A,g)^{-1}z_A\).  Hence
\[
  \mathcal A_Y(A^g)
  =
  (-1)^{p(A)}z_{A^g}\psi_{A^g}
  =
  (-1)^{p(A)}z_A\psi_A
  =
  \mathcal A_Y(A).
\]
For two fillings, the ratio \(z_{0,A}/z_{1,A}\) is a scalar because both
vectors live in the same graded dual line.  The common supertrace sign
is not a gauge-cocycle factor; it belongs to the closed-interface comparison
of \(Z_{\rm DF}(Y_0)\) with the orientation-reversed vector from
\(-Y_1\).  With the graded orientation-reversal convention in
\eqref{eq:graded-orientation-reversal-dual}, the gluing law
\eqref{eq:dai-freed-boundary-gluing-law} identifies this comparison with the
closed Dai--Freed phase of \(Y_0\cup_M(-Y_1)\).  Pairing both numerator and
denominator with the same boundary vector \(\Psi(A)\) then gives
\eqref{eq:dai-freed-filling-change-closed-phase}; for an odd interface, the
minus sign is exactly the supertrace sign in
\eqref{eq:dai-freed-supertrace-interface}.
\end{proof}

\paragraph{Composition through an intermediate boundary.}
The same line-valued mechanism controls the anomaly factor when a physical
spacetime is cut into successive pieces.  We use the ordinary bordism
orientation convention of Chapter~\ref{ch:tqft-bordism-functoriality-extended}:
a bordism \(Y_{01}:M_0\to M_1\) has
\[
  \partial Y_{01}=(-M_0)\sqcup M_1 .
\]
Orientation reversal dualizes the anomaly line,
\(L_{-M}\simeq L_M^\vee\), in the graded sense of
\eqref{eq:graded-orientation-reversal-dual}.  Since the Dai--Freed object
for a manifold with boundary is a vector in the inverse line over the full
boundary, this gives
\[
  Z_{\rm DF}(Y_{01})
  \in
  L_{\partial Y_{01}}^\vee
  \simeq
  (L_{-M_0}\otimes L_{M_1})^\vee
  \simeq
  L_{M_0}\otimes L_{M_1}^\vee .
\]
The last display fixes the source-to-target tensor order used below; the
graded symmetry needed to pass from the formal dual of a tensor product to
this order is included in that identification and is not repeated as a
separate sign in later formulas.
Thus the Dai--Freed bulk has the dual variance to a relative boundary-QFT
transition amplitude from \(M_0\) to \(M_1\), which would live in
\(L_{M_0}^\vee\otimes L_{M_1}\).  The former is the inverse anomaly factor
that cancels the latter; it is not itself the physical transition kernel.

For successive invertible bulk bordisms \(Y_{01}:M_0\to M_1\) and
\(Y_{12}:M_1\to M_2\), the resulting Dai--Freed vectors have the form
\[
  Z_{\rm DF}(Y_{01})\in L_{M_0}\otimes L_{M_1}^\vee,
  \qquad
  Z_{\rm DF}(Y_{12})\in L_{M_1}\otimes L_{M_2}^\vee.
\]
Their composition is obtained by taking the supertrace over the two
\(M_1\)-factors in the displayed order:
\begin{equation}
  Z_{\rm DF}(Y_{12}\circ Y_{01})
  =
  \bigl(\operatorname{id}_{L_{M_0}}
  \otimes\operatorname{str}_{L_{M_1}}
  \otimes\operatorname{id}_{L_{M_2}^\vee}\bigr)
  \bigl(Z_{\rm DF}(Y_{01})\otimes Z_{\rm DF}(Y_{12})\bigr).
  \label{eq:dai-freed-interface-composition}
\end{equation}
Here
\(\operatorname{str}_{L_{M_1}}:L_{M_1}^\vee\otimes L_{M_1}\to\mathbb C\)
is the Dai--Freed supertrace \eqref{eq:dai-freed-supertrace-interface}.  If
\(p(M_1)=0\), this is ordinary evaluation.  If \(p(M_1)=1\), it is minus
ordinary evaluation.  The formulas above are written in the order in which
the intermediate factors are evaluated; any other reordering of odd
determinant or Pfaffian factors must include the Koszul sign rule.

This is only the anomaly-line component of locality.  A boundary
Hilbert-space matrix element, or a Euclidean transition amplitude, also
requires an actual state space, pairing or integration measure, gauge
treatment, and topology/convergence statement.  Once that physical gluing
operation has been separately constructed, the anomaly factors on the common
cut are paired away by \eqref{eq:dai-freed-interface-composition}; the line
identity alone does not construct the sum or integral over physical degrees
of freedom.

\begin{proposition}[Intermediate-cut frame cancellation]
\label{prop:dai-freed-intermediate-cut-frame-cancellation}
Choose local frames \(e_i\) of the anomaly lines \(L_{M_i}\).  Write
\[
  Z_{\rm DF}(Y_{01})=z_{01}\,e_0\otimes e_1^\vee,
  \qquad
  Z_{\rm DF}(Y_{12})=z_{12}\,e_1\otimes e_2^\vee.
\]
Then the composed coefficient in
\(L_{M_0}\otimes L_{M_2}^\vee\) is
\((-1)^{p(M_1)}z_{01}z_{12}\), and it is independent of the choice of
frame \(e_1\) on the glued boundary.  If the frame on
\(M_1\) is changed by \(e_1\mapsto q e_1\), \(q\in U(1)\), the two
coefficients transform as \(z_{01}\mapsto qz_{01}\) and
\(z_{12}\mapsto q^{-1}z_{12}\); their product is unchanged.  The same statement
holds for real Pfaffian lines with \(q=\pm1\), with
\((-1)^{p_{\rm Pf}(M_1)}\) in place of the complex determinant parity.
\end{proposition}

\begin{proof}
The graded trace satisfies
\(\operatorname{str}_{L_{M_1}}(e_1^\vee\otimes e_1)=(-1)^{p(M_1)}\).
Substituting the two displayed expressions into
\eqref{eq:dai-freed-interface-composition} gives
\[
  Z_{\rm DF}(Y_{12}\circ Y_{01})
  =
  (-1)^{p(M_1)}z_{01}z_{12}\,e_0\otimes e_2^\vee.
\]
After the frame change \(e_1'=q e_1\), the same vectors are written as
\[
  Z_{\rm DF}(Y_{01})=(qz_{01})\,e_0\otimes (e_1')^\vee,
  \qquad
  Z_{\rm DF}(Y_{12})=(q^{-1}z_{12})\,e_1'\otimes e_2^\vee.
\]
The product of the two coefficients is still \(z_{01}z_{12}\), and the
supertrace sign \((-1)^{p(M_1)}\) is intrinsic to the intermediate
superline, not to the chosen frame.  Hence the coefficient of the glued
bordism in the remaining endpoint superlines is frame independent.

For three successive pieces \(Y_{01},Y_{12},Y_{23}\), with endpoint frames
kept in the fixed order
\[
  e_0\otimes e_1^\vee\otimes e_1\otimes e_2^\vee
  \otimes e_2\otimes e_3^\vee,
\]
the two possible parenthesizations give
\begin{align*}
  &\bigl(Z_{\rm DF}(Y_{23})\circ Z_{\rm DF}(Y_{12})\bigr)\circ
  Z_{\rm DF}(Y_{01}) \\
  &\qquad =
  Z_{\rm DF}(Y_{23})\circ
  \bigl(Z_{\rm DF}(Y_{12})\circ Z_{\rm DF}(Y_{01})\bigr) \\
  &\qquad =
  (-1)^{p(M_1)+p(M_2)}z_{01}z_{12}z_{23}\,
  e_0\otimes e_3^\vee .
\end{align*}
Thus associativity is ordinary associativity of scalar multiplication plus
the two graded traces on the two common interfaces.  When the glued
interfaces all have even index, these signs are \(+1\) and the formula
reduces to the ungraded determinant-line calculation.
\end{proof}

When a relative boundary-theory amplitude
\(\mathcal U_{02}\in L_{M_0}^\vee\otimes L_{M_2}\) is paired with the
composed bulk vector, all exposed anomaly lines are evaluated and the result
is an ordinary number.  If the physical theory supplies a valid cut formula
for \(\mathcal U_{02}\) through \(M_1\), then
\eqref{eq:dai-freed-interface-composition} says that the anomaly-line part of
that cut formula is independent of the artificial frame on \(M_1\).  All
phase data are carried by the exposed endpoint lines and by the
closed-manifold eta phase that compares two different global anomaly-line
trivializations, as in \eqref{eq:dai-freed-filling-change-closed-phase}.

\paragraph{Local descent as the infinitesimal limit of inflow.}
For a contractible loop of background fields, the Dai--Freed holonomy reduces
to the exponential of the descent integral determined by the local anomaly
polynomial.  This statement is not an additional index theorem.  Once the
Bismut--Freed curvature formula and the adiabatic holonomy formula have been
fixed with the conventions above, the remaining passage from global inflow to
the local consistent anomaly is ordinary Stokes transgression in
background-field space.

Let \(\mathcal B\) be a coordinate patch in the space of structured
backgrounds and let \(\gamma=\partial\Sigma\) be an oriented contractible loop
bounding a smooth two-chain \(\Sigma\subset\mathcal B\).  Write
\(\Omega_{\mathrm{BF}}\) for the normalized curvature two-form of the
Bismut--Freed connection.  For the bounding circle spin structure induced
from the disk, and with the determinant-line rather than inverse-line
variance in \eqref{eq:bismut-freed-holonomy}, a disk contained in this patch
satisfies
\begin{equation}
  \operatorname{Hol}_{\gamma}(\operatorname{Det}D)
  =
  \exp\!\left(
    2\pi\ii\int_{\Sigma}\Omega_{\mathrm{BF}}
  \right).
  \label{eq:bf-contractible-loop-stokes}
\end{equation}
The curvature theorem identifies \(\Omega_{\mathrm{BF}}\) with the fiber
integral over \(M\) of the degree-\((d+2)\) component of
\(\widehat A(TM)\operatorname{ch}(E)\), again in the same sign convention.
Equivalently, if \(I_{d+2}\) denotes the normalized anomaly polynomial on
the universal family over \(\mathcal B\times M\), then
\begin{equation}
  \int_\Sigma \Omega_{\mathrm{BF}}
  =
  \int_{\Sigma\times M} I_{d+2}.
  \label{eq:bf-curvature-family-index-integral}
\end{equation}

Now specialize one tangent direction of \(\Sigma\) to a gauge-orbit
direction generated by a Lie-algebra-valued parameter \(\lambda\).  Choose a
Chern--Simons representative \(I_{d+1}^{(0)}\) on physical spacetime and its
descent form \(I_d^{(1)}(\lambda,A)\), so that
\begin{equation}
  I_{d+2}=d I_{d+1}^{(0)},
  \qquad
  \delta_\lambda I_{d+1}^{(0)}
  =
  d I_d^{(1)}(\lambda,A).
  \label{eq:local-descent-from-family-curvature}
\end{equation}
Stokes on the two-parameter cylinder in background space turns
\(\int_{\Sigma\times M}I_{d+2}\) into the integral of
\(I_d^{(1)}\) over the physical boundary slice.  Thus the infinitesimal
gauge variation of a local trivialization of the fermion determinant line is
\begin{equation}
  \delta_\lambda\log Z_{\mathrm{fermion}}(A)
  =
  2\pi\ii\int_M I_d^{(1)}(\lambda,A),
  \label{eq:local-descent-from-dai-freed}
\end{equation}
up to the sign choice already fixed by the determinant-line convention in
\eqref{eq:bismut-freed-holonomy}.  This is the same local descent class used
in the perturbative chiral-anomaly calculation and in the anomaly-inflow
construction developed earlier in the monograph.
Local counterterms change \(I_d^{(1)}\) by a relative coboundary; they change
the local connection form on the determinant line but not its curvature or
its contractible-loop holonomy.

The qualification ``contractible'' is essential.  If the loop is not filled
by a disk in the background groupoid, equations
\eqref{eq:bf-contractible-loop-stokes}--\eqref{eq:local-descent-from-dai-freed}
do not determine the holonomy.  The reduced eta invariant may then contain a
flat anomaly-line contribution invisible to the local polynomial.  The
Witten \(SU(2)\) anomaly above is exactly of this type: the local cubic
descent vanishes, while a nontrivial mapping-torus mod-two index detects the
large-gauge holonomy.

\paragraph{Theorem boundary for the global-anomaly construction.}
The anomaly statements in this chapter use four external global-analysis
inputs, and the QFT conclusions are the consequences derived from them in the
fixed conventions above.  The APS theorem supplies the Fredholm boundary
problem and the identity
\[
  \operatorname{Ind}(D_X^{+,\mathrm{APS}})
  =
  \int_X[\widehat A(TX)\operatorname{ch}(E)]-\xi(B_{\partial X})
\]
with the product-collar and reduced-eta conventions fixed in
\eqref{eq:aps-collar-dirac-form}--\eqref{eq:aps-index-formula}.  From this
input the text derives the cylinder congruence, the endpoint-kernel term, the
orientation sign in the spectral-flow bookkeeping, and the local descent
variation of the eta phase.  The Bismut--Freed theorem supplies the Quillen
determinant line, its connection, its local curvature, and the mapping-torus
holonomy formula.  From these data the text derives the quotient-groupoid
cocycle criterion: a chiral determinant descends to the quotient by gauge
transformations if and only if the anomaly line admits an equivariant
trivialization; local counterterms alter the cocycle by a coboundary; based
loop holonomies are invariant.  The Dai--Freed construction supplies the
boundary vector in the inverse anomaly line, its orientation-dual composition
law, and the cancellation of anomaly factors on artificial cuts.  From those
line-valued statements the text derives the inflow cancellation criterion and
the contractible-loop reduction to the same descent representative used in
the perturbative anomaly chapter.  They do not by themselves construct the
physical Hilbert-space pairing, path-integral measure, gauge fixing, or
convergence needed for a full QFT gluing theorem.  Finally, the real mod-two
index theorem
supplies the value of the Pfaffian holonomy for the Witten \(SU(2)\)
mapping torus.  The trace-delta normalization, the \(2j\equiv1\pmod4\)
parity criterion, the tensor-product cancellation condition, and the
separation between local cubic anomaly and global sign anomaly are then
finite representation-theoretic and anomaly-line consequences in the
monograph's conventions.

Thus the remaining quoted material is pure global-analysis infrastructure:
Fredholmness and heat-kernel local index theory with APS boundary
conditions, construction and holonomy of the determinant/Pfaffian line, and
the real \(KO\)-theoretic mod-two index calculation on the Witten mapping
torus.  The QFT-facing anomaly claims are the groupoid descent, local
descent, inflow, and cancellation statements derived from those inputs in the
preceding paragraphs.

\section{Interacting Theories and Anomaly Lines}

For free fermions, the anomaly line is constructed from a family of Dirac
operators.  For an interacting QFT, the same structure should be formulated
without choosing quadratic fields.  The background groupoid has objects given
by structured background fields and morphisms given by gauge transformations,
diffeomorphisms, and identifications of background data.  An anomaly line is
a functor from this groupoid to the groupoid of complex lines, equipped with
compatible gluing data.  The partition function of the anomalous theory is a
section of this functorial line.

Local counterterms change the trivialization of the line.  They shift the
connection one-form and the cocycle representative, but they do not change
the curvature class or the holonomy class.  Thus a genuine global anomaly is
not removed by changing local coordinates on the space of actions.

\begin{openproblem}[Global anomaly line for interacting QFT]
\label{op:global-anomaly-line-interacting-qft}
Construct the anomaly line of an interacting QFT over the space of background
fields, prove its local curvature and global holonomy formulae, and identify
the conditions under which it is represented by an invertible bulk field
theory.  The free-fermion determinant and Pfaffian lines developed above are
models for this structure, not a general construction of it.
\end{openproblem}
