\chapter{Microlocal Spectrum Condition And Hadamard Geometry}
\label{ch:curved-microlocal-spectrum-condition}

\ConventionsMixed

This chapter develops the microlocal form of the positive-frequency condition
on curved spacetimes before the volume uses Hadamard states, Wick products,
time-ordered products, and renormalized stress tensors.  The replacement for
translation-invariant spectral support is a wavefront-set condition on
distributions, and the local geometric form of the two-point function is the
Hadamard parametrix.

\section{Wavefront Set}

Let \(u\in\mathcal D'(X)\) be a distribution on a smooth manifold \(X\).  A
point \((x,k)\in T^*X\setminus\{0\}\) is absent from the wavefront set
\(\operatorname{WF}(u)\) if there is a cutoff \(\chi\in C_c^\infty(X)\) with
\(\chi(x)\ne0\) and a conic neighborhood \(\Gamma\) of \(k\) such that the
Fourier transform of \(\chi u\) decays faster than every power in \(\Gamma\).
The wavefront set records both the singular support and the covector
directions of oscillation.

\begin{definition}[Wavefront set in a chart]
\label{def:wavefront-set-chart}
Let \(u\in\mathcal D'(X)\).  In a coordinate chart \(U\subset X\), write
\(\widehat{\chi u}(\xi)\) for the Fourier transform of the compactly
supported distribution \(\chi u\), \(\chi\in C_c^\infty(U)\), in the chosen
coordinates.  A nonzero covector \(k\in T_x^*X\) is a regular direction for
\(u\) if there exist \(\chi\in C_c^\infty(U)\), \(\chi(x)\ne0\), and an open
conic neighborhood \(\Gamma\subset\mathbb R^D\setminus\{0\}\) of the
coordinate representative of \(k\) such that, for every \(N\),
\begin{equation}
  |\widehat{\chi u}(\xi)|
  \leq C_N(1+|\xi|)^{-N},
  \qquad
  \xi\in\Gamma .
  \label{eq:microlocal-rapid-decay-direction}
\end{equation}
The wavefront set \(\operatorname{WF}(u)\) is the complement, in
\(T^*X\setminus0\), of the regular directions.
\end{definition}

The definition is independent of the chart.  Under a change of coordinates,
the localized Fourier transform is an oscillatory integral whose phase has
non-degenerate mixed Hessian at the point under consideration; integration
by parts in directions where the transformed covector is separated from the
old singular cone preserves rapid decay.  Thus \(\operatorname{WF}(u)\) is a
closed conic subset of \(T^*X\setminus0\).

\begin{theorem}[Product criterion for distributions]
\label{thm:microlocal-product-criterion}
Let \(u,v\in\mathcal D'(X)\).  Suppose that there is no
\begin{equation}
  (x,k)\in\operatorname{WF}(u)
  \quad\hbox{and}\quad
  (x,-k)\in\operatorname{WF}(v).
  \label{eq:microlocal-no-opposite-covectors}
\end{equation}
Then the pointwise product \(uv\) has a unique distributional extension
which agrees with the ordinary product when one factor is smooth.  Moreover
\begin{equation}
\begin{aligned}
  \operatorname{WF}(uv)
  \subset&
  \operatorname{WF}(u)\cup\operatorname{WF}(v)\\
  &\cup
  \{(x,k+\ell):(x,k)\in\operatorname{WF}(u),\
        (x,\ell)\in\operatorname{WF}(v),\ k+\ell\ne0\}.
\end{aligned}
  \label{eq:microlocal-product-wavefront-bound}
\end{equation}
\end{theorem}

\begin{proof}
The statement is local, so work in one coordinate chart and multiply by a
cutoff equal to \(1\) near the point.  If \(u\) and \(v\) are compactly
supported distributions, the formal Fourier transform of the product is the
convolution
\begin{equation}
  \widehat{uv}(\xi)
  =
  (2\pi)^{-D}\int \widehat u(\eta)\widehat v(\xi-\eta)\,\dd\eta .
  \label{eq:microlocal-product-convolution}
\end{equation}
The obstruction to defining this integral distributionally at large
\(|\eta|\) is simultaneous non-decay of \(\widehat u(\eta)\) and
\(\widehat v(\xi-\eta)\).  As \(|\eta|\to\infty\), the second direction is
asymptotic to \(-\eta\).  Thus a large-frequency obstruction can occur only
from an opposite pair of singular covectors \((k,-k)\), which is excluded by
\eqref{eq:microlocal-no-opposite-covectors}.

Choose a conic partition of unity in the \(\eta\)-variable.  On pieces where
\(\eta\) avoids \(\operatorname{WF}(u)\), the factor \(\widehat u(\eta)\) is
rapidly decreasing after localization.  On pieces where \(\xi-\eta\) avoids
\(\operatorname{WF}(v)\), the factor \(\widehat v(\xi-\eta)\) is rapidly
decreasing.  The exclusion of opposite cones gives such a finite conic cover
of the large-\(\eta\) region.  The bounded-\(\eta\) region is harmless
because Fourier transforms of compactly supported distributions have at most
polynomial growth.  This defines
\eqref{eq:microlocal-product-convolution}, and the same conic decomposition
proves rapid decay of \(\widehat{uv}\) away from the three sets displayed in
\eqref{eq:microlocal-product-wavefront-bound}.

Uniqueness follows by regularizing \(u\) and \(v\) with a smoothing
approximate identity.  The estimates above show convergence in every allowed
conic region and independence of the regularization.  If one factor is
smooth, the construction reduces to ordinary multiplication because the
smooth factor has rapidly decreasing localized Fourier transform.
\end{proof}

This criterion is the local analytic reason that Wick products and
time-ordered products require subtraction before coincident points are taken.

\section{Hadamard Two-Point Condition}

Let \(M\) be a globally hyperbolic Lorentzian spacetime and let \(\phi\) be a
real scalar field satisfying
\begin{equation}
  P_M\phi=0,\qquad P_M=-\nabla^\mu\nabla_\mu+m^2+\xi R[g] .
  \label{eq:microlocal-kg-operator}
\end{equation}
A quasifree state is determined by its two-point function
\[
  \omega_2(x,y)=\omega(\phi(x)\phi(y)).
\]
The time orientation on tangent vectors induces a convention on covectors:
a nonzero causal covector \(k\in T_x^*M\) is future directed when its metric
dual
\[
  k^\sharp=g^{\mu\nu}k_\nu\partial_\mu
\]
is a future-directed causal vector.  In mostly-plus Minkowski coordinates,
this means that a future null covector has \(k_0<0\).  This convention makes
the positive-frequency plane wave \(e^{-\ii Et+\ii{\bf p}\cdot{\bf x}}\)
carry future null covector \((-E,{\bf p})\).

The microlocal Hadamard condition is
\[
  \operatorname{WF}(\omega_2)
  =
  \{(x,k_x;y,-k_y): (x,k_x)\sim(y,k_y),\ k_x\in\overline V_+^*\}.
\]
Here \((x,k_x)\sim(y,k_y)\) means that \(x\) and \(y\) are joined by a null
geodesic and that \(k_y\) is the parallel transport of \(k_x\) along that
geodesic.  The cone \(\overline V_+^*\) is the closed future causal cone in
the cotangent space.

The minus sign on the second covector is not a convention that can be
dropped.  It records that a two-point distribution is a distribution on
\(M\times M\).  If the same future covector is transported from \(x\) to
\(y\), the corresponding singular phase is locally of the form
\[
  \exp\bigl(\ii\lambda(\Phi(x)-\Phi(y))\bigr),
  \qquad \lambda\to+\infty,
\]
so the covector on the second copy of \(M\) is the negative of the
transported covector.

\section{Higher Microlocal Spectrum Condition}

For non-Gaussian states, the two-point condition is not enough.  The
curved-spacetime replacement of the Wightman spectral condition is a family
of cone constraints on all \(n\)-point distributions.

\begin{definition}[Microlocal spectrum cone]
\label{def:microlocal-spectrum-cone}
Let \(x_1,\ldots,x_n\in M\).  A covector tuple
\[
  (x_1,k_1;\ldots;x_n,k_n)\in T^*M^n\setminus0
\]
belongs to \(\Gamma_n(M)\) if there is a finite directed graph with vertex
set \(\{1,\ldots,n\}\) such that each edge \(e:i\to j\) is assigned a
future-directed causal covector field \(p_e\) parallel transported along a
future-directed causal curve from \(x_i\) to \(x_j\), and
\begin{equation}
  k_i
  =
  \sum_{e:i\to j}p_e(x_i)
  -
  \sum_{e:j\to i}p_e(x_i)
  \qquad (i=1,\ldots,n).
  \label{eq:microlocal-spectrum-graph-balance}
\end{equation}
The empty graph gives the zero tuple, which is omitted from
\(T^*M^n\setminus0\).
\end{definition}

\begin{definition}[Microlocal spectrum condition for a state]
\label{def:microlocal-spectrum-condition-state}
A state \(\omega\) on a scalar field algebra satisfies the microlocal
spectrum condition when all \(n\)-point distributions
\[
  \omega_n(f_1,\ldots,f_n)
  =
  \omega(\phi(f_1)\cdots\phi(f_n))
\]
obey
\begin{equation}
  \operatorname{WF}(\omega_n)\subset\Gamma_n(M).
  \label{eq:microlocal-spectrum-condition-all-n}
\end{equation}
\end{definition}

For \(n=2\), a graph with one edge \(1\to2\) gives
\[
  (k_1,k_2)=(p,-p),
\]
which is the sign pattern in the Hadamard two-point condition.  For
quasifree Hadamard states, Wick's theorem reduces every \(n\)-point
distribution to a finite sum of products of two-point functions.  Applying
Theorem~\ref{thm:microlocal-product-criterion} and the graph description
above shows that the microlocal spectrum condition for all \(n\) follows
from the two-point Hadamard condition.

\section{Hadamard Parametrix}

The microlocal condition specifies the singular covector directions.  The
Hadamard parametrix specifies the local distribution with those singularities.
This section constructs its coefficients from the operator
\eqref{eq:microlocal-kg-operator}.  The construction is local in a normal
convex neighborhood \(\mathcal U\), so every pair \(x,y\in\mathcal U\) is
joined by a unique geodesic lying in \(\mathcal U\).

\begin{definition}[Synge function and boundary value]
\label{def:microlocal-synge-boundary-value}
Let \(\sigma(x,y)\) be Synge's world function, one half of the signed squared
geodesic distance from \(y\) to \(x\).  Choose a smooth time function \(T\) on
\(\mathcal U\) and set
\begin{equation}
  \sigma_\epsilon(x,y)
  =
  \sigma(x,y)+\ii\epsilon(T(x)-T(y))+\epsilon^2,
  \qquad \epsilon>0 .
  \label{eq:microlocal-sigma-epsilon}
\end{equation}
All formulas involving negative powers or logarithms of \(\sigma_\epsilon\)
are understood as distributional boundary values as \(\epsilon\downarrow0\).
\end{definition}

The basic coincidence identities are
\begin{equation}
  \sigma^{;\mu}\sigma_{;\mu}=2\sigma,\qquad
  [\sigma]=0,\qquad
  [\sigma_{;\mu}]=0,\qquad
  [\sigma_{;\mu\nu}]=g_{\mu\nu},\qquad
  [\Box_g\sigma]=4 ,
  \label{eq:microlocal-synge-coincidence}
\end{equation}
where brackets denote the coincidence limit \(y\to x\).  The dimension four
enters only through the last identity.

\begin{definition}[Van Vleck determinant]
\label{def:microlocal-van-vleck}
The van Vleck determinant is the biscalar density ratio
\begin{equation}
  \Delta(x,y)
  =
  \frac{
  \det\!\bigl(-\nabla_\mu^x\nabla_\nu^y\sigma(x,y)\bigr)
  }{
  |g(x)|^{1/2}|g(y)|^{1/2}
  },
  \label{eq:microlocal-van-vleck}
\end{equation}
written in any pair of local coordinates and interpreted invariantly by the
density factors.  It satisfies \(\Delta(x,x)=1\).
\end{definition}

\begin{lemma}[Hadamard coefficient \(U\)]
\label{lem:microlocal-u-transport}
In four spacetime dimensions, the coefficient of \(\sigma_\epsilon^{-1}\) in
the Hadamard parametrix is
\begin{equation}
  U(x,y)=\Delta(x,y)^{1/2}.
  \label{eq:microlocal-u-delta}
\end{equation}
Equivalently, \(U\) is the unique smooth biscalar along each geodesic from
\(y\) to \(x\) satisfying
\begin{equation}
  2\sigma^{;\mu}\nabla_\mu U+(\Box_g\sigma-4)U=0,
  \qquad U(x,x)=1 .
  \label{eq:microlocal-u-transport}
\end{equation}
\end{lemma}

\begin{proof}
For a smooth biscalar \(U\), use
\(\sigma^{;\mu}\sigma_{;\mu}=2\sigma\) to compute
\begin{equation}
  \Box_g(\sigma^{-1})=(4-\Box_g\sigma)\sigma^{-2}.
  \label{eq:microlocal-box-sigma-inverse}
\end{equation}
Since \(P_M=-\Box_g+m^2+\xi R[g]\), the \(\sigma^{-2}\) term in
\(P_{M,x}(U/\sigma)\) is
\begin{equation}
  \frac{
  2\sigma^{;\mu}\nabla_\mu U+(\Box_g\sigma-4)U
  }{\sigma^2}.
  \label{eq:microlocal-u-singular-coefficient}
\end{equation}
The zeroth-order term \(m^2+\xi R[g]\) contributes only order
\(\sigma^{-1}\).  Therefore the most singular coefficient vanishes exactly
when \eqref{eq:microlocal-u-transport} holds.

It remains to identify the solution.  Along the geodesic parameterized by
\(s\), the operator \(\sigma^{;\mu}\nabla_\mu\) is the radial derivative
weighted by squared geodesic distance.  The Jacobi-field equation for the
geodesic exponential map gives
\begin{equation}
  \sigma^{;\mu}\nabla_\mu\log\Delta
  =
  4-\Box_g\sigma .
  \label{eq:microlocal-delta-transport}
\end{equation}
Taking one half of this equation gives
\eqref{eq:microlocal-u-transport} for \(U=\Delta^{1/2}\), and the coincidence
normalization follows from the normal-coordinate expansion of the exponential
map.  Uniqueness follows from the first-order ordinary differential equation
along the geodesic.
\end{proof}

\begin{definition}[Hadamard parametrix in four dimensions]
\label{def:microlocal-hadamard-parametrix}
Let
\begin{equation}
  V(x,y)\sim\sum_{j\geq0}v_j(x,y)\sigma(x,y)^j
  \label{eq:microlocal-v-expansion}
\end{equation}
be a formal expansion by smooth biscalars.  The local Hadamard parametrix is
\begin{equation}
  H_{\epsilon,\ell}(x,y)
  =
  \frac1{8\pi^2}
  \left[
    \frac{U(x,y)}{\sigma_\epsilon(x,y)}
    +
    V(x,y)\log\!\left(\frac{\sigma_\epsilon(x,y)}{\ell^2}\right)
  \right],
  \qquad \ell>0,
  \label{eq:microlocal-hadamard-parametrix}
\end{equation}
where \(U\) is fixed by Lemma~\ref{lem:microlocal-u-transport} and the
jet of \(V\) at the diagonal is fixed by the recursion below.
\end{definition}

\begin{theorem}[Hadamard coefficient recursion]
\label{thm:microlocal-hadamard-recursion}
With \(P_M\) as in \eqref{eq:microlocal-kg-operator}, the logarithmic
coefficient is determined recursively by
\begin{equation}
  \bigl(2\sigma^{;\mu}\nabla_\mu+\Box_g\sigma-2\bigr)v_0
  =
  P_{M,x}U ,
  \label{eq:microlocal-v0-transport}
\end{equation}
and, for \(j\geq0\),
\begin{equation}
  \bigl(2\sigma^{;\mu}\nabla_\mu+\Box_g\sigma+2j\bigr)v_{j+1}
  =
  \frac1{j+1}P_{M,x}v_j .
  \label{eq:microlocal-vj-transport}
\end{equation}
At coincidence,
\begin{equation}
  v_0(x,x)
  =
  \frac12\left(m^2+\left(\xi-\frac16\right)R[g](x)\right),
  \label{eq:microlocal-v0-coincidence}
\end{equation}
and
\begin{equation}
  v_{j+1}(x,x)
  =
  \frac{1}{2(j+1)(j+2)}
  \left.P_{M,x}v_j(x,y)\right|_{y=x}.
  \label{eq:microlocal-vj-coincidence}
\end{equation}
For every finite \(N\), the partial parametrix built from
\(\sum_{j=0}^{N}v_j\sigma^j\) satisfies \(P_{M,x}H_{\epsilon,\ell}^{[N]}\)
smooth up to the order fixed by the omitted coefficient; a smooth \(V\) with
the prescribed diagonal jet exists by Borel's lemma, and changing that smooth
representative changes \(H_{\epsilon,\ell}\) only by a smooth biscalar.
\end{theorem}

\begin{proof}
The calculation is performed for \(x\neq y\) before the boundary value is
taken.  Put \(s=\sigma(x,y)\).  For a smooth biscalar \(v\),
\begin{equation}
  P_{M,x}(v\log s)
  =
  (P_{M,x}v)\log s
  -
  \frac{
  2\sigma^{;\mu}\nabla_\mu v+(\Box_g\sigma-2)v
  }{s}.
  \label{eq:microlocal-p-v-log}
\end{equation}
Together with Lemma~\ref{lem:microlocal-u-transport}, the remaining
\(s^{-1}\) term in \(P_{M,x}(U/s+v_0\log s)\) is canceled precisely by
\eqref{eq:microlocal-v0-transport}.

For \(j\geq0\), the coefficient of \(s^j\log s\) in
\[
  P_{M,x}\left(\sum_{\ell\geq0}v_\ell s^\ell\log s\right)
\]
has two contributions.  The term \(v_j s^j\) contributes
\[
  s^jP_{M,x}v_j .
\]
The lower-derivative part of \(P_{M,x}(v_{j+1}s^{j+1})\) contributes
\begin{equation}
  -(j+1)s^j
  \left(
  2\sigma^{;\mu}\nabla_\mu v_{j+1}
  +(\Box_g\sigma+2j)v_{j+1}
  \right).
  \label{eq:microlocal-log-coefficient-step}
\end{equation}
Vanishing of this coefficient gives
\eqref{eq:microlocal-vj-transport}.  Taking \(y\to x\) in
\eqref{eq:microlocal-v0-transport} and using
\(\sigma^{;\mu}=0\), \(\Box_g\sigma=4\), \(U(x,x)=1\), and
\begin{equation}
  \left.\Box_{g,x}U(x,y)\right|_{y=x}=\frac16R[g](x)
  \label{eq:microlocal-box-u-coincidence}
\end{equation}
gives \eqref{eq:microlocal-v0-coincidence}.  Taking the coincidence limit of
\eqref{eq:microlocal-vj-transport} gives
\eqref{eq:microlocal-vj-coincidence}.

At each step the transport operator has nonzero diagonal coefficient, so the
diagonal value and then all derivatives transverse to the diagonal are fixed
by differentiating the transport equation.  Borel's lemma produces a smooth
biscalar with the prescribed infinite jet.  Two such choices differ by a
smooth biscalar flat at the diagonal, hence by a smooth contribution to
\eqref{eq:microlocal-hadamard-parametrix}.
\end{proof}

\begin{remark}[Hadamard form of a quasifree two-point function]
\label{rem:microlocal-hadamard-smooth-remainder}
A quasifree two-point function \(\omega_2\) is locally Hadamard in
\(\mathcal U\) precisely when
\begin{equation}
  \omega_2(x,y)-H_{\epsilon,\ell}(x,y)
  \label{eq:microlocal-hadamard-smooth-remainder}
\end{equation}
extends to a smooth biscalar on \(\mathcal U\times\mathcal U\), after taking
the distributional boundary value.  Changing \(\ell\) changes the subtraction
by the smooth local term
\begin{equation}
  -\frac{1}{4\pi^2}V(x,y)\log(\ell'/\ell).
  \label{eq:microlocal-hadamard-scale-change}
\end{equation}
\end{remark}

The singular terms in \eqref{eq:microlocal-hadamard-parametrix} are fixed by
the local equation \(P_M\omega_2=0\) modulo smooth terms and by the
positive-frequency boundary value in
\eqref{eq:microlocal-sigma-epsilon}.  Theorem
\ref{thm:microlocal-hadamard-recursion} constructs the unique diagonal jet
needed to remove the singular part generated by applying \(P_M\) in either
variable.  Therefore any two local parametrices with the same \(\ell\) differ
smoothly.  A state is in the Hadamard class exactly when its two-point
function has this universal singular part and no further singular remainder,
which is \eqref{eq:microlocal-hadamard-smooth-remainder}.  The scale formula
follows by subtracting the two logarithms in
\eqref{eq:microlocal-hadamard-parametrix}.

\section{Propagation Of Singularities}

For a real principal type operator \(P\), singularities of solutions to
\(Pu=0\) propagate along the Hamiltonian flow of the principal symbol.  For
the Klein--Gordon operator \eqref{eq:microlocal-kg-operator}, the principal
symbol is \(-g^{\mu\nu}k_\mu k_\nu\).  The characteristic set is unchanged if
one uses the metric Hamiltonian
\[
  p(x,k)=g^{\mu\nu}(x)k_\mu k_\nu .
\]
The two Hamiltonian vector fields differ by an overall sign, hence by a
reversal of the bicharacteristic parameter, not by a different null geodesic.
The Hamiltonian vector field on \(T^*M\) is
\[
  H_p
  =
  \frac{\partial p}{\partial k_\mu}\frac{\partial}{\partial x^\mu}
  -
  \frac{\partial p}{\partial x^\mu}\frac{\partial}{\partial k_\mu}.
\]
Thus its integral curves satisfy
\begin{equation}
  \dot x^\mu=2g^{\mu\nu}(x)k_\nu,
  \qquad
  \dot k_\mu=-(\partial_\mu g^{\alpha\beta})(x)k_\alpha k_\beta .
  \label{eq:microlocal-kg-hamilton-equations}
\end{equation}
Let \(v^\mu=\dot x^\mu\).  Then
\[
  v^\mu=2k^{\sharp\mu},
  \qquad
  g_{\mu\nu}v^\mu v^\nu=4p(x,k).
\]
On \(p=0\), the projected curve \(x(s)\) is therefore null.  Lowering the
index in the first Hamilton equation gives \(k_\mu=\frac12g_{\mu\nu}v^\nu\).
Differentiate this identity and use the second Hamilton equation together
with
\[
  \partial_\mu g^{\alpha\beta}
  =
  -g^{\alpha\rho}g^{\beta\sigma}\partial_\mu g_{\rho\sigma}.
\]
One obtains
\[
  g_{\mu\nu}\left(\ddot x^\nu+\Gamma^\nu_{\rho\sigma}
       \dot x^\rho\dot x^\sigma\right)=0 .
\]
Since \(g\) is non-degenerate, \(x(s)\) is an affinely parametrized null
geodesic, and \(k\) is parallel transported along it up to the fixed factor
\(1/2\).

Hence once the two-point function has positive-frequency wavefront
directions on one Cauchy surface, the equation of motion propagates those
directions along null geodesics.  This is the geometric content of the
microlocal spectrum condition.

\section{Wick Products}

For a Hadamard state, define a Wick-square coordinate by point splitting:
\[
  :\!\phi^2\!:_H(f)
  =
  \int_M f(x)
  \lim_{y\to x}
  \bigl(\phi(x)\phi(y)-H(x,y)\mathbf 1\bigr)\,\dd\operatorname{vol}_g(x).
\]
This formula separates two operations which are often compressed in notation.
If \(H'=H+w\), with \(w\) smooth near the diagonal, then
\[
  :\!\phi^2\!:_{H'}(f)
  =
  :\!\phi^2\!:_{H}(f)
  -
  \int_M f(x)w(x,x)\,\dd\operatorname{vol}_g(x)\,\mathbf 1 .
\]
The diagonal \(w(x,x)\) is a smooth function.  For a generic change of
Hadamard coordinate it need not be a local curvature polynomial, and under
the corresponding algebra isomorphism a transported state gives the same
expectation value for the same abstract observable.

A finite local Wick-renormalization change is a stricter comparison: it
compares two locally covariant Wick prescriptions in a fixed physical state.
Under the usual covariance, scaling-dimension, and field-equation conditions,
the four-dimensional scalar Wick square has the finite prescription freedom
\[
  :\!\phi^2\!:'_H
  =
  :\!\phi^2\!:_H
  +
  c_1R+c_2m^2 .
\]
Thus the microlocal fact is the smooth diagonal shift \(w(x,x)\); the
curvature polynomial is the additional local-covariant prescription freedom.
Chapter~\ref{ch:point-splitting-stress-tensor-renormalization} gives the
point-splitting version of this distinction, and
Chapter~\ref{ch:curved-perturbative-algebraic-qft} tracks the same
distinction through Hadamard-coordinate transport of interacting Wick
polynomials.

\section{Time-Ordered Products}

The time-ordered two-point distribution has a wavefront set with both
future and past directions arranged according to time ordering.  Products of
time-ordered distributions at coincident points violate the wavefront-set
product criterion.  Perturbative algebraic renormalization extends these
distributions from configuration space with diagonals removed to the full
configuration space.  The extension freedom is local and covariant, with
degree bounded by scaling.  Contact terms in curved spacetime are precisely
the local distributions supported on these diagonals, with covariance and
Ward identities imposed as constraints.

\begin{openproblem}[Interacting microlocal spectrum condition]
\label{op:interacting-microlocal-spectrum-condition}
Construct interacting locally covariant QFTs on curved backgrounds with
Hadamard states, local Wick and time-ordered products, stress tensor,
background gauge fields, and anomaly lines, and prove the microlocal spectrum
condition for the resulting nonperturbative Wightman-type distributions.
\end{openproblem}
