\chapter{Witten-Donaldson Theory And The Seiberg-Witten Comparison}
\label{ch:tqft-witten-donaldson-seiberg-witten-comparison}

\ConventionsEuclidean

This chapter develops the four-dimensional cohomological gauge theory whose
correlation functions are Donaldson invariants and explains the precise
mathematical content of the comparison with Seiberg-Witten monopole theory.
There are three logically distinct layers.  The first layer is finite
dimensional differential geometry: anti-self-dual connections, monopole
equations, deformation complexes, orientations, compactifications, and
intersection pairings.  The second layer is cohomological field theory:
twisted supersymmetry, descent, \(Q\)-exact metric variation, and contact
terms.  The third layer is the four-dimensional RG statement: the twisted
non-Abelian ultraviolet theory has a Coulomb-branch Wilsonian description
whose singular fibers are Abelian monopole theories.  The chapter keeps these
layers separate because they have different hypotheses and different
mathematical status.

\paragraph{How the comparison argument is organized.}
The local mechanisms in this chapter should be read as a dependency ladder,
not as interchangeable derivations of the final comparison formula.  The
Donaldson moduli-space layer defines the ASD invariants, the \(\mu\)-classes,
simple type, compactification faces, and wall links.  The cohomological
field-theory layer explains why these finite-dimensional objects are the
localization targets of a twisted gauge theory, and why metric changes produce
BV boundary terms rather than arbitrary chamber jumps.  The Coulomb-branch
layer supplies a different description of the same putative four-dimensional
QFT: a regular \(u\)-plane integral, contact terms, electromagnetic-duality
determinant lines, and singular-fiber monopole theories.  The comparison
statement at the end is the assertion that these three layers can be glued by
one \(Q\)-compatible Wilsonian construction.  Consequently a wall-link
identity, a monopole dimension count, a \(u\)-plane boundary model, or a
blow-up coefficient table is only a test of one arrow in the ladder; none of
them alone is a proof of the Donaldson--Seiberg-Witten comparison.

\section{Four-Manifold And Gauge Conventions}

Let \(X\) be a closed connected oriented smooth four-manifold equipped with a
Riemannian metric \(g\).  The Hodge star operator satisfies
\[
  *_g^2=1\quad\text{on }\Omega^2(X),
\]
and the bundle of two-forms decomposes as
\[
  \Lambda^2T^*X=\Lambda^{2,+}T^*X\oplus\Lambda^{2,-}T^*X .
\]
For a two-form \(\omega\), write
\[
  \omega^\pm=\frac{1}{2}(\omega\pm *_g\omega).
\]
The integer \(b_2^+(X)\) is the dimension of a maximal positive subspace of
the intersection form on \(H^2(X;\mathbb R)\), equivalently the dimension of
the space of harmonic self-dual two-forms for any metric \(g\).  The Euler
characteristic and signature are
\[
  \chi(X)=2-2b_1(X)+b_2^+(X)+b_2^-(X),
  \qquad
  \sigma(X)=b_2^+(X)-b_2^-(X).
\]

Let \(P\to X\) be a principal \(SU(2)\) bundle and let
\(\operatorname{ad}P=P\times_{\operatorname{Ad}}\mathfrak{su}(2)\).  The
gauge group is
\[
  \mathcal G_P=\Gamma(X,P\times_{\operatorname{Ad}}SU(2)).
\]
The monograph uses the gauge trace convention fixed in the gauge-theory
volumes:
\[
  \operatorname{tr}(t^at^b)=\delta^{ab}
\]
for the fundamental \(SU(N)\) generators.  With this convention the Euclidean
Yang--Mills action is written
\[
  S_{\mathrm{YM}}(A)
  =
  \frac{1}{4g_{\mathrm{YM}}^2}
  \int_X d\operatorname{vol}_g\,
  \operatorname{tr}(F_{A,\mu\nu}F_A^{\mu\nu})
  -\frac{i\theta}{8\pi^2}\int_X\operatorname{tr}(F_A\wedge F_A).
\]
Equivalently,
\[
  S_{\mathrm{YM}}(A)
  =
  \frac{1}{2g_{\mathrm{YM}}^2}
  \int_X\operatorname{tr}(F_A\wedge *_gF_A)
  -\frac{i\theta}{8\pi^2}\int_X\operatorname{tr}(F_A\wedge F_A).
\]
The instanton number is
\[
  k(P)
  =
  -\frac{1}{8\pi^2}\int_X\operatorname{tr}(F_A\wedge F_A).
\]
It is independent of the connection \(A\).  For \(SU(2)\), this integer is
the second Chern number \(c_2(P)\) in the present trace convention.

The curvature decomposes as \(F_A=F_A^+ + F_A^-\).  Define
\[
  \|F_A^\pm\|^2=\int_X\operatorname{tr}(F_A^\pm\wedge *_gF_A^\pm).
\]
Then
\[
  \int_X\operatorname{tr}(F_A\wedge *_gF_A)
  =
  \|F_A^+\|^2+\|F_A^-\|^2,
  \qquad
  \int_X\operatorname{tr}(F_A\wedge F_A)
  =
  \|F_A^+\|^2-\|F_A^-\|^2.
\]
Hence
\[
  \frac{1}{2g_{\mathrm{YM}}^2}
  \int_X\operatorname{tr}(F_A\wedge *_gF_A)
  =
  \frac{1}{g_{\mathrm{YM}}^2}\|F_A^+\|^2
  +\frac{4\pi^2}{g_{\mathrm{YM}}^2}k(P).
\]
The anti-self-dual equation
\[
  F_A^+=0
\]
therefore saturates the topological bound in the sector \(k(P)\ge 0\).  The
normalization \(4\pi^2 k/g_{\mathrm{YM}}^2\) is the trace-delta version of
the common half-trace expression \(8\pi^2 k/g_{\mathrm{ht}}^2\), with
\(g_{\mathrm{ht}}^2=2g_{\mathrm{YM}}^2\).

\section{Donaldson-Witten Cohomological Fields}

The Donaldson twist of the pure four-dimensional \(\mathcal N=2\) vector
multiplet was defined in
Chapter~\ref{ch:tqft-twists-of-supersymmetric-theories}.  After twisting,
the fields are organized as differential forms on \(X\):
\[
  A\in\mathcal A(P),\qquad
  \phi,\bar\phi\in\Omega^0(X,\operatorname{ad}P),\qquad
  \psi\in\Omega^1(X,\operatorname{ad}P),
\]
\[
  \chi^+\in\Omega^{2,+}(X,\operatorname{ad}P),\qquad
  \eta\in\Omega^0(X,\operatorname{ad}P),
\]
together with an auxiliary self-dual field
\[
  H^+\in\Omega^{2,+}(X,\operatorname{ad}P).
\]
The scalar odd symmetry \(Q\) acts by
\[
\begin{aligned}
  &QA=\psi,\qquad &&Q\psi=-d_A\phi,\qquad &&Q\phi=0,\\
  &Q\bar\phi=\eta,\qquad &&Q\eta=[\phi,\bar\phi],\qquad
  &&Q\chi^+=H^+,\qquad QH^+=[\phi,\chi^+] .
\end{aligned}
\]
Thus \(Q^2\) is the infinitesimal gauge transformation with parameter
\(-\phi\) on \(A\) and with the corresponding adjoint action on all adjoint
fields.  On gauge-invariant observables, \(Q^2=0\).

The bosonic localization term may be chosen so that its \(H^+\)-dependent part
has the form
\[
  \int_X\operatorname{tr}\left(
  \frac{1}{2}H^+\wedge *_gH^+
  +H^+\wedge *_gF_A^+
  \right).
\]
The algebraic equation of motion for \(H^+\) gives \(H^+=-F_A^+\), and the
corresponding bosonic square is proportional to
\[
  \|F_A^+\|^2 .
\]
Hence the weak-coupling localization locus in a fixed instanton sector is the
anti-self-dual equation \(F_A^+=0\).  Metric independence of \(Q\)-closed
correlators follows from the Ward identity of
Remark~\ref{rem:twist-metric-independence} once a regulator, contact
prescription, and compactification boundary prescription preserving the
required \(Q\)-relations have been fixed.

\section{The Anti-Self-Dual Moduli Space}

The anti-self-dual moduli space in the sector \(P\) is
\[
  \mathcal M(P,g)
  =
  \{A\in\mathcal A(P)\mid F_A^+=0\}/\mathcal G_P .
\]
A connection \(A\) is irreducible when its stabilizer in \(\mathcal G_P\) is
the center of \(SU(2)\).  At an irreducible anti-self-dual connection, the
linearized equation and infinitesimal gauge transformations form the elliptic
complex
\[
  0\longrightarrow
  \Omega^0(X,\operatorname{ad}P)
  \xrightarrow{\,d_A\,}
  \Omega^1(X,\operatorname{ad}P)
  \xrightarrow{\,d_A^+\,}
  \Omega^{2,+}(X,\operatorname{ad}P)
  \longrightarrow0 .
\]
Its cohomology groups are denoted
\[
  H_A^0,\qquad H_A^1,\qquad H_A^2 .
\]
Irreducibility gives \(H_A^0=0\).  The Zariski tangent space is \(H_A^1\), and
\(H_A^2\) is the obstruction space.  When \(H_A^2=0\), the moduli space is a
smooth manifold near \([A]\) of dimension \(\dim H_A^1\).

The gauge-fixed elliptic operator is
\[
  D_A=d_A^*\oplus d_A^+:
  \Omega^1(X,\operatorname{ad}P)
  \longrightarrow
  \Omega^0(X,\operatorname{ad}P)\oplus
  \Omega^{2,+}(X,\operatorname{ad}P).
\]
Its index equals the virtual dimension
\[
  \dim_{\mathrm{vir}}\mathcal M(P,g)
  =
  \dim H_A^1-\dim H_A^0-\dim H_A^2 .
\]

\paragraph{Index of the anti-self-dual complex.}
For \(P\to X\) a principal \(SU(2)\) bundle with instanton number \(k\), the
virtual dimension of the anti-self-dual moduli space is
\[
  \dim_{\mathrm{vir}}\mathcal M(P,g)
  =
  8k-3(1-b_1(X)+b_2^+(X)).
\]

The Atiyah--Singer index theorem applied to
\[
  D_A=d_A^*\oplus d_A^+
\]
gives
\[
  \operatorname{ind}D_A
  =
  -2p_1(\operatorname{ad}P)
  -\frac{\dim SU(2)}{2}\bigl(\chi(X)+\sigma(X)\bigr).
\]
For a principal \(SU(2)\) bundle,
\[
  p_1(\operatorname{ad}P)=-4c_2(P)=-4k .
\]
Also
\[
  \frac{\chi(X)+\sigma(X)}{2}
  =
  1-b_1(X)+b_2^+(X).
\]
Since \(\dim SU(2)=3\), substitution gives
\[
  \operatorname{ind}D_A
  =
  8k-3(1-b_1(X)+b_2^+(X)).
\]
This index is \(\dim H_A^1-\dim H_A^0-\dim H_A^2\), the virtual dimension of
the elliptic complex.

The determinant line of the complex is
\[
  \mathcal L_A
  =
  \det H_A^1\otimes(\det H_A^0)^{-1}\otimes(\det H_A^2)^{-1}.
\]
An orientation of the instanton moduli problem is an orientation of this line
over the irreducible moduli space.  For simply connected \(X\), such an
orientation is fixed by an orientation of \(H^2_+(X;\mathbb R)\).  For general
\(X\), the homology orientation is an orientation of
\[
  H^1(X;\mathbb R)\oplus H^2_+(X;\mathbb R).
\]

The Uhlenbeck compactification \(\overline{\mathcal M}(P,g)\) adds ideal
instantons:
\[
  ([A'],x_1,\ldots,x_\ell),
  \qquad
  k(P')+\ell=k(P).
\]
Geometrically, curvature concentration at finitely many points carries the
missing second Chern number.  Donaldson invariants are defined by extending
the relevant cohomology classes to this compactification or to an equivalent
smooth compactified model, with reducible strata and wall crossing treated
separately when they occur.

\section{Donaldson Series, Simple Type, And Uhlenbeck Boundary Terms}

For the comparison formula it is useful to use the standard \(SO(3)\) or
\(U(2)\) notation.  Fix an integral lift \(w\in H^2(X;\mathbb Z)\) of the
second Stiefel--Whitney class of the adjoint bundle; the earlier \(SU(2)\)
case is the special case \(w=0\).  Let \(x\in H_0(X;\mathbb Z)\) denote the
positive generator represented by a point and let \(h\in H_2(X;\mathbb R)\).
The Donaldson generating function is the degree-selected formal series
\begin{equation}
  D_X^w(p,h)
  =
  \sum_{r,n\ge0}
  \frac{p^r}{r!\,n!}
  D_X^w(x^rh^n),
  \label{eq:donaldson-generating-function}
\end{equation}
where terms whose total Donaldson degree does not equal the relevant virtual
dimension are zero.  The point class has Donaldson degree four and a surface
class has Donaldson degree two.

\begin{definition}[Donaldson simple type and Donaldson series]
\label{def:donaldson-simple-type-series}
Assume \(b_1(X)=0\) and \(b_2^+(X)>1\), so that the Donaldson invariants are
independent of the metric chamber after the orientation convention has been
fixed.  The pair \((X,w)\) has Donaldson simple type if
\begin{equation}
  D_X^w\bigl((x^2-4)z\bigr)=0
  \label{eq:donaldson-simple-type-relation}
\end{equation}
for every homogeneous element \(z\) of the Donaldson homology algebra.  In
that case the reduced Donaldson series is
\begin{equation}
  \mathbb D_X^w(h)
  =
  D_X^w\left(\left(1+\frac{x}{2}\right)\exp(h)\right).
  \label{eq:reduced-donaldson-series}
\end{equation}
The normalization \(x^2=4\) in
\eqref{eq:donaldson-simple-type-relation} is part of the Donaldson-invariant
convention used in this chapter; changing the normalization of the point
observable rescales the numerical constant but not the structure of the
comparison.
\end{definition}

\begin{quotedtheorem}[Donaldson simple-type structure theorem]
\label{qt:donaldson-simple-type-structure}
Let \(X\) be a closed connected oriented smooth four-manifold with
\(b_1(X)=0\) and \(b_2^+(X)>1\).  If \((X,w)\) has Donaldson simple type, then
there is a finite set
\[
  \mathcal B_D(X,w)\subset H^2(X;\mathbb Z)
\]
and rational coefficients \(a_K^w\) such that
\begin{equation}
  \mathbb D_X^w(h)
  =
  \exp\!\left(\frac{1}{2}Q_X(h,h)\right)
  \sum_{K\in\mathcal B_D(X,w)}
  a_K^w\,\exp\!\bigl(\langle K,h\rangle\bigr).
  \label{eq:donaldson-simple-type-exponential-form}
\end{equation}
The classes \(K\) are characteristic:
\begin{equation}
  K\equiv w_2(TX)\pmod 2.
  \label{eq:donaldson-basic-characteristic}
\end{equation}
They are called Donaldson basic classes.  The theorem is a theorem of the
finite-dimensional Donaldson-invariant construction, not a theorem that the
four-dimensional twisted QFT has been constructed non-perturbatively.
\end{quotedtheorem}

\paragraph{What the structure theorem supplies.}
The theorem contains two different kinds of information, and they should not be
confused.  The first is geometric.  Donaldson invariants define a linear
functional on the graded algebra generated by the point class \(x\) and
homology classes of \(X\).  The simple-type relation
\eqref{eq:donaldson-simple-type-relation} says that the point operator has
minimal polynomial \(T^2-4\) on this functional.  Passing to the reduced series
\eqref{eq:reduced-donaldson-series} selects the \(+2\)-eigensummand in the
normalization used here.  The structure theorem then proves, using the
finite-dimensional Donaldson moduli-space construction, gluing, blow-up, and
compactification control, that the remaining surface-source dependence has
only finitely many characteristic exponential weights after a universal
quadratic factor is removed.  The finiteness of the set
\(\mathcal B_D(X,w)\), the characteristic parity
\eqref{eq:donaldson-basic-characteristic}, and the fact that no polynomial
prefactors multiply the exponentials are part of this deep theorem.

The second kind of information is algebraic and follows once the finite set is
known.  Put
\begin{equation}
  F_X^w(h)
  =
  \exp\!\left(-\frac{1}{2}Q_X(h,h)\right)\mathbb D_X^w(h).
  \label{eq:donaldson-reduced-finite-transform}
\end{equation}
Then \eqref{eq:donaldson-simple-type-exponential-form} is equivalent to saying
that \(F_X^w\) is a finite exponential sum on the vector space
\(H_2(X;\mathbb R)\).  For every \(v\in H_2(X;\mathbb R)\), it obeys the
constant-coefficient ordinary differential equation
\begin{equation}
  \prod_{K\in\mathcal B_D(X,w)}
  \left(
    \frac{d}{dt}-\langle K,v\rangle
  \right)
  F_X^w(h+tv)
  =
  0 .
  \label{eq:donaldson-finite-exponential-ode}
\end{equation}
Conversely, a finite-dimensional space of translation derivatives with
simultaneously diagonal characters \(K\) gives a sum of exponentials
\(\exp\langle K,h\rangle\).  This converse is linear algebra over the
commuting derivative operators; the hard Donaldson input is the geometric
statement that the derivative representation is finite and semisimple with
characteristic weights.

The exponential form in
\eqref{eq:donaldson-simple-type-exponential-form} is not a slogan.  It says
that all dependence on the ordinary intersection form is the universal
Gaussian factor \(\exp(Q_X/2)\), while all differentiable information in the
simple-type sector is carried by finitely many characteristic linear
functionals \(K\).  A coefficient \(a_K^w\) is defined only after the
orientation, lift \(w\), point-observable normalization, and compactification
extension of the \(\mu\)-classes have been fixed.

\begin{definition}[Donaldson simple-type normalization convention]
\label{def:donaldson-simple-type-normalization-convention}
The symbol \(a_K^w\) in
\eqref{eq:donaldson-simple-type-exponential-form} is used only after the
following data have been declared.
\[
\begin{array}{ll}
  \text{homology orientation} &
  \text{an orientation of }H^1(X;\mathbb R)\oplus H^{2,+}(X;\mathbb R),\\[1mm]
  \text{lift} &
  w\in H^2(X;\mathbb Z),\quad w\equiv w_2(\operatorname{ad}P)\pmod2,\\[1mm]
  \text{point class} &
  x\in H_0(X;\mathbb Z)\text{ normalized by }D_X^w((x^2-4)z)=0,\\[1mm]
  \text{surface source} &
  h\in H_2(X;\mathbb R)\text{ paired with }K\in H^2(X;\mathbb Z),\\[1mm]
  \text{compactification} &
  \text{an extension of the relevant }\mu\text{-classes to the boundary.}
\end{array}
\]
Changing any one of these data changes the numerical coefficients by a
controlled convention: reversing the homology orientation reverses all
oriented counts, changing \(w\) changes the chamber and phase convention of
the \(SO(3)\) or \(U(2)\) invariant, and rescaling the point observable changes
the constant \(4\) in \eqref{eq:donaldson-simple-type-relation}.
\end{definition}

\begin{remark}[Coefficient extraction from the reduced series]
\label{rem:donaldson-basic-coefficient-extraction}
Assume the hypotheses and conventions of
Theorem~\ref{qt:donaldson-simple-type-structure}.  Define
\[
  F_X^w(h)
  =
  \exp\!\left(-\frac{1}{2}Q_X(h,h)\right)\mathbb D_X^w(h).
\]
Then
\begin{equation}
  F_X^w(h)=\sum_{K\in\mathcal B_D(X,w)}a_K^w e^{\langle K,h\rangle}.
  \label{eq:donaldson-finite-exponential-sum}
\end{equation}
For any \(v\in H_2(X;\mathbb R)\) and \(m\ge0\),
\begin{equation}
  \left.\frac{d^m}{dt^m}\right|_{t=0}F_X^w(tv)
  =
  \sum_{K\in\mathcal B_D(X,w)}
  a_K^w\,\langle K,v\rangle^m.
  \label{eq:donaldson-moment-extraction}
\end{equation}
Consequently, after the finite set \(\mathcal B_D(X,w)\) is known, the
coefficients \(a_K^w\) are determined by finitely many Donaldson invariants:
choose vectors \(v_1,\ldots,v_N\) such that the numbers
\(\langle K,v_i\rangle\) separate the basic classes, and invert the resulting
finite Vandermonde system.

Multiplying \eqref{eq:donaldson-simple-type-exponential-form} by
\(\exp(-Q_X/2)\) gives
\eqref{eq:donaldson-finite-exponential-sum}.  Differentiating the finite sum
in the direction \(v\) gives \eqref{eq:donaldson-moment-extraction}.  If the
linear functionals \(K\) are separated by the chosen directions, the matrix of
moments is a product of ordinary Vandermonde matrices.  Its determinant is
nonzero precisely because the exponential weights have distinct evaluations.
Thus the \(a_K^w\) are not extra formal symbols: they are the finite Fourier
coefficients of the reduced Donaldson series in the basis of characteristic
exponentials.
\end{remark}

\paragraph{Uhlenbeck boundary as BV boundary data.}
Let \(\mathcal M_k^{\rm sm}\) be a smooth irreducible ASD stratum in
instanton number \(k\), and let
\begin{equation}
  \overline{\mathcal M}_k^{\,U}
  =
  \mathcal M_k^{\rm sm}
  \sqcup
  \bigsqcup_{\ell\ge1}
  \mathcal M_{k-\ell}^{\rm sm}\times{\rm Sym}^{\ell}(X)
  \sqcup
  \{\text{reducible and obstructed strata}\}
  \label{eq:donaldson-uhlenbeck-strata}
\end{equation}
denote the relevant part of the Uhlenbeck compactification.  If the
Donaldson-Witten path integral is represented as a BV pushforward over
\(\mathcal M_k^{\rm sm}\), then the compactification strata contribute
boundary functionals:
\begin{equation}
  \Delta_{\rm res}(\pi_*^{\mathcal M_k^{\rm sm}}\rho)
  =
  -\sum_{\ell\ge1}
  \mathcal B_{\mathcal M_{k-\ell}^{\rm sm}\times{\rm Sym}^{\ell}(X)}(\rho)
  -\mathcal B_{\rm red/obs}(\rho).
  \label{eq:donaldson-uhlenbeck-bv-boundary}
\end{equation}
Consequently a Donaldson invariant obtained from the compactified moduli
space is a QME-preserving cohomological correlator only after the boundary
terms in \eqref{eq:donaldson-uhlenbeck-bv-boundary} have been cancelled by
codimension, by an extension of the insertion classes, by a chamber
wall-crossing term, or by an explicitly supplied compactification datum.

This is the singular-stratum obstruction formula
\eqref{eq:singular-stratum-localized-qme-obstruction} applied to the
finite-dimensional localization locus of Donaldson-Witten theory.  The strata
\(\mathcal M_{k-\ell}^{\rm sm}\times{\rm Sym}^{\ell}(X)\) are the
small-instanton boundary strata: \(\ell\) units of second Chern number have
left the smooth connection and are recorded as unordered points of \(X\).
Reducible and obstructed strata are additional singular strata of the same
BV-boundary type.  Removing collar neighborhoods of all these strata gives a
manifold with boundary; BV Stokes produces exactly the right hand side of
\eqref{eq:donaldson-uhlenbeck-bv-boundary}.  If the sum vanishes or is
cancelled, the residual semidensity satisfies the QME; otherwise the
formal localization expression is not yet a cohomological invariant.

\section{Donaldson Observables And The \texorpdfstring{\(\mu\)}{mu}-Map}

The basic degree-four gauge-invariant polynomial in the twisted scalar is
\[
  \mathcal O^{(0)}
  =
  \frac{1}{8\pi^2}\operatorname{tr}\phi^2 .
\]
With the \(Q\)-rules above, the first descendants may be chosen as
\[
  \mathcal O^{(1)}
  =
  \frac{1}{4\pi^2}\operatorname{tr}(\phi\psi),
\]
\[
  \mathcal O^{(2)}
  =
  \frac{1}{4\pi^2}\operatorname{tr}
  \left(\phi F_A+\frac{1}{2}\psi\wedge\psi\right),
\]
and the sequence continues to four-forms.  They obey the descent equations
\[
  d\mathcal O^{(j)}
  =
  \{Q,\mathcal O^{(j+1)}\}_{\mathrm{gr}},
  \qquad j=0,1,2,3 .
\]
For a cycle \(\gamma_j\in H_j(X;\mathbb Z)\), define
\[
  I(\gamma_j)=\int_{\gamma_j}\mathcal O^{(j)} .
\]
The \(Q\)-cohomological degree of \(I(\gamma_j)\) is \(4-j\).  In the finite
dimensional instanton description this insertion represents the Donaldson
\(\mu\)-class
\[
  \mu(\gamma_j)\in
  H^{4-j}(\mathcal M(P,g);\mathbb Q).
\]
If \(\mathbb P\to X\times\mathcal M(P,g)\) denotes a universal bundle over a
smooth irreducible locus, then the class can be written as
\[
  \mu(\gamma_j)
  =
  -\frac{1}{4}
  p_1(\operatorname{ad}\mathbb P)/\gamma_j ,
\]
where the slash denotes slant product with \(\gamma_j\).  This formula is a
differential-geometric version of the descent construction.  In particular,
for a point \(x\in X\), \(\mu(x)\) has degree four, and for a surface
\(\Sigma\subset X\), \(\mu(\Sigma)\) has degree two.

Given homology classes \(\gamma_{j_1},\ldots,\gamma_{j_r}\), the Donaldson
correlator in the smooth compact case is the pairing
\[
  \left\langle
    I(\gamma_{j_1})\cdots I(\gamma_{j_r})
  \right\rangle_{P,g}
  =
  \int_{\overline{\mathcal M}(P,g)}
  \mu(\gamma_{j_1})\smile\cdots\smile\mu(\gamma_{j_r})
\]
when
\[
  \sum_{\alpha=1}^r(4-j_\alpha)
  =
  \dim_{\mathrm{vir}}\mathcal M(P,g).
\]
If the total degree differs, the correlator vanishes after the usual
degree-selection rule, apart from contact-term and reducible-wall phenomena
specified by the compactified theory.

\section{Spin\texorpdfstring{\(^c\)}{c} Structures And Monopole Equations}

A \(\operatorname{Spin}^c\) structure \(\mathfrak s\) on \(X\) consists of
rank-two Hermitian bundles \(S^\pm_{\mathfrak s}\), a determinant line
\[
  L_{\mathfrak s}=\det S^+_{\mathfrak s}=\det S^-_{\mathfrak s},
\]
and Clifford multiplication
\[
  \rho:T^*X\longrightarrow
  \operatorname{Hom}(S^+_{\mathfrak s},S^-_{\mathfrak s})
\]
compatible with the metric.  The characteristic class
\[
  c_1(\mathfrak s)=c_1(L_{\mathfrak s})\in H^2(X;\mathbb Z)
\]
is characteristic:
\[
  c_1(\mathfrak s)\equiv w_2(TX)\pmod2 .
\]

\begin{proposition}[\(\operatorname{Spin}^c\) structures as characteristic lifts]
\label{prop:spinc-characteristic-lifts}
On an oriented smooth four-manifold \(X\), the set
\({\rm Spin}^c(X)\) of isomorphism classes of
\(\operatorname{Spin}^c\) structures is a torsor for \(H^2(X;\mathbb Z)\).
Tensoring \(\mathfrak s\) by a Hermitian line bundle \(M\to X\) gives a new
\(\operatorname{Spin}^c\) structure \(\mathfrak s\otimes M\) with
\begin{equation}
  c_1(\mathfrak s\otimes M)
  =
  c_1(\mathfrak s)+2c_1(M).
  \label{eq:spinc-c1-tensor-line}
\end{equation}
The image of the map
\[
  c_1:{\rm Spin}^c(X)\longrightarrow H^2(X;\mathbb Z)
\]
is the affine set of integral characteristic classes
\begin{equation}
  \{c\in H^2(X;\mathbb Z)\mid c\equiv w_2(TX)\pmod2\}.
  \label{eq:spinc-characteristic-image}
\end{equation}
If \(H^2(X;\mathbb Z)\) has two-torsion, distinct
\(\operatorname{Spin}^c\) structures can have the same first Chern class;
the structure, not only \(c_1\), is then part of the data.
\end{proposition}

\begin{proof}
Existence of \(\operatorname{Spin}^c\) structures on oriented smooth
four-manifolds is the topological fact \(W_3(TX)=\beta(w_2(TX))=0\); this is
equivalent to the existence of an integral characteristic lift of \(w_2(TX)\).
The group \(\operatorname{Spin}^c(4)\) is
\[
  \operatorname{Spin}^c(4)
  =
  (\operatorname{Spin}(4)\times U(1))/\{\pm(1,1)\}.
\]
A \(\operatorname{Spin}^c\) structure lifting the oriented frame bundle can
therefore be tensored by any principal \(U(1)\)-bundle, equivalently by a
Hermitian line bundle \(M\).  This action is free and transitive because the
difference between two lifts of the same oriented frame bundle is exactly a
\(U(1)\)-bundle.  The determinant representation squares the additional
\(U(1)\) factor, so tensoring by \(M\) changes the determinant line by
\[
  L_{\mathfrak s\otimes M}
  =
  L_{\mathfrak s}\otimes M^{\otimes2},
\]
which proves \eqref{eq:spinc-c1-tensor-line}.  The reduction of
\(c_1(L_{\mathfrak s})\) modulo two is the obstruction \(w_2(TX)\) to a
spin lift, so every \(c_1(\mathfrak s)\) is characteristic.  Conversely, if
\(c\) is another integral lift of \(w_2(TX)\), then
\(c-c_1(\mathfrak s)\) lies in the kernel of reduction modulo two.  The
exact sequence associated to
\[
  0\longrightarrow \mathbb Z
  \xrightarrow{\times2}
  \mathbb Z
  \longrightarrow \mathbb Z_2
  \longrightarrow0
\]
identifies this kernel with \(2H^2(X;\mathbb Z)\), so
\(c=c_1(\mathfrak s)+2c_1(M)\) for some line bundle \(M\).  If
\(2c_1(M)=0\), tensoring by \(M\) leaves the first Chern class unchanged,
which is the stated two-torsion caveat.
\end{proof}

For a unitary connection \(B\) on \(L_{\mathfrak s}\), the Dirac operator is
\[
  \not D_B:\Gamma(S^+_{\mathfrak s})\longrightarrow
  \Gamma(S^-_{\mathfrak s}).
\]
Let \(q\in\Gamma(S^+_{\mathfrak s})\).  The quadratic map
\[
  \sigma:S^+_{\mathfrak s}\longrightarrow\Lambda^{2,+}T^*X\otimes i\mathbb R
\]
is defined by
\[
  \rho(\sigma(q))=(q q^\dagger)_0 ,
\]
where \((q q^\dagger)_0\) is the traceless Hermitian endomorphism of
\(S^+_{\mathfrak s}\) obtained from the rank-one endomorphism
\(v\mapsto q\,\langle q,v\rangle\).  This equation fixes the normalization of
\(\sigma\) used below.

The unperturbed Seiberg-Witten equations are
\[
  \not D_B q=0,
  \qquad
  F_B^+=\sigma(q).
\]
A perturbation by a self-dual imaginary two-form \(\eta^+\) gives
\[
  \not D_B q=0,
  \qquad
  F_B^+=\sigma(q)+\eta^+ .
\]
The gauge group is
\[
  \mathcal G_L=C^\infty(X,U(1)),
\]
acting by
\[
  u\cdot(B,q)=(B-u^{-1}du,\,uq).
\]
The moduli space is
\[
  \mathcal M_{\mathrm{SW}}(\mathfrak s,g,\eta^+)
  =
  \{(B,q)\mid \not D_Bq=0,\ F_B^+=\sigma(q)+\eta^+\}/\mathcal G_L .
\]

The linearized, gauge-fixed elliptic operator at a solution is
\[
  \mathcal D_{(B,q)}:
  \Omega^1(X;i\mathbb R)\oplus\Gamma(S^+_{\mathfrak s})
  \longrightarrow
  \Omega^0(X;i\mathbb R)\oplus
  \Omega^{2,+}(X;i\mathbb R)\oplus\Gamma(S^-_{\mathfrak s}) .
\]
Its index is the expected dimension of the monopole moduli space.

\paragraph{Index of the monopole deformation problem.}
For a \(\operatorname{Spin}^c\) structure \(\mathfrak s\), the expected
dimension of the Seiberg-Witten moduli space is
\[
  d_{\mathrm{SW}}(\mathfrak s)
  =
  \frac{c_1(\mathfrak s)^2-\bigl(2\chi(X)+3\sigma(X)\bigr)}{4}.
\]

The real index of the Dirac part is twice the complex index:
\[
  2\,\operatorname{ind}_{\mathbb C}\not D_B
  =
  2\int_X \widehat A(TX)e^{c_1(\mathfrak s)/2}
  =
  \frac{c_1(\mathfrak s)^2-\sigma(X)}{4}.
\]
The Abelian gauge-fixing and self-dual curvature operator
\[
  d^*\oplus d^+:\Omega^1(X;i\mathbb R)
  \longrightarrow
  \Omega^0(X;i\mathbb R)\oplus\Omega^{2,+}(X;i\mathbb R)
\]
has index
\[
  -(1-b_1(X)+b_2^+(X)).
\]
Adding the two real indices gives
\[
  \frac{c_1(\mathfrak s)^2-\sigma(X)}{4}
  -(1-b_1(X)+b_2^+(X)).
\]
The identity
\[
  2\chi(X)+3\sigma(X)
  =
  4(1-b_1(X)+b_2^+(X))+\sigma(X)
\]
converts this expression to
\[
  \frac{c_1(\mathfrak s)^2-\bigl(2\chi(X)+3\sigma(X)\bigr)}{4}.
\]

\section{Compactness Input From The Weitzenbock Identity}

The monopole equations have compactness properties stronger than the
instanton equations because the spinor obeys a nonlinear elliptic equation
with a quartic term.  The Dirac Weitzenbock identity is
\[
  \not D_B^*\not D_B q
  =
  \nabla_B^*\nabla_B q
  +\frac{s_g}{4}q
  +\frac{1}{2}\rho(F_B^+)q ,
\]
where \(s_g\) is the scalar curvature.  Pairing with \(q\), integrating over
\(X\), and using \(\not D_Bq=0\) gives
\[
  0=
  \int_X\left(
    |\nabla_Bq|^2
    +\frac{s_g}{4}|q|^2
    +\frac{1}{2}\langle\rho(F_B^+)q,q\rangle
  \right)d\operatorname{vol}_g .
\]
With the normalization
\(\rho(\sigma(q))=(q q^\dagger)_0\), the unperturbed curvature equation gives
\[
  \frac{1}{2}\langle\rho(F_B^+)q,q\rangle
  =
  \frac{1}{4}|q|^4 .
\]
Thus solutions obey
\[
  \int_X\left(
    |\nabla_Bq|^2
    +\frac{s_g}{4}|q|^2
    +\frac{1}{4}|q|^4
  \right)d\operatorname{vol}_g
  =
  0
\]
in the unperturbed case.  With perturbation \(\eta^+\), the right side is
controlled by the \(L^2\)-norm of \(\eta^+\) and the negative part of scalar
curvature.  Elliptic bootstrapping then bounds \(q\) and \(F_B^+\), while the
topological identity fixing \(c_1(L_{\mathfrak s})\) controls the anti-self-dual
component modulo gauge.  This is the analytic source of compactness for the
monopole moduli space after excluding reducible solutions by a generic
perturbation when the usual chamber hypotheses allow it.

\section{Seiberg-Witten Invariants And Basic Classes}

Assume \(b_2^+(X)>1\), choose a homology orientation, and choose generic
metric/perturbation data so that the monopole moduli spaces are cut out
transversely and contain no reducible solutions in the relevant classes.  If
\(\mathfrak s\) has expected dimension \(d_{\rm SW}(\mathfrak s)=0\), its
Seiberg-Witten invariant is the signed count
\begin{equation}
  {\rm SW}_X(\mathfrak s)
  =
  \#_{\rm or}\,\mathcal M_{\rm SW}(\mathfrak s,g,\eta^+).
  \label{eq:sw-zero-dimensional-count}
\end{equation}
For positive expected dimension, the invariant pairs the fundamental cycle
of the compact oriented monopole moduli space with the standard cohomology
classes obtained from the universal \(U(1)\) bundle and, when \(b_1(X)>0\),
from \(H_1(X)\).  In this chapter the zero-dimensional case is the main one
because it is the simple-type sector appearing in Witten's comparison.

\begin{definition}[Seiberg-Witten basic class and simple type]
\label{def:sw-basic-class-simple-type}
An integral class \(K\in H^2(X;\mathbb Z)\) is a Seiberg-Witten basic class
if \(K=c_1(\mathfrak s)\) for a \(\operatorname{Spin}^c\) structure
\(\mathfrak s\) whose Seiberg-Witten invariant is nonzero.  The four-manifold
has Seiberg-Witten simple type if every basic class satisfies
\begin{equation}
  d_{\rm SW}(\mathfrak s)=0,
  \qquad\text{equivalently}\qquad
  K^2=2\chi(X)+3\sigma(X).
  \label{eq:sw-simple-type-condition}
\end{equation}
\end{definition}

The adjective ``basic'' is not a synonym for ``allowed.''  Allowed
\(\operatorname{Spin}^c\) structures are the characteristic lifts of
Proposition~\ref{prop:spinc-characteristic-lifts}.  Basic classes are the
much smaller subset detected by a nonzero monopole count or by the
higher-dimensional monopole pairings.  The Weitzenbock estimate gives the
compactness input for each fixed characteristic class; the finiteness of
nonzero basic classes is a global theorem of the Seiberg-Witten invariant
construction.

\begin{example}[K3, elliptic surfaces, and blow-up bookkeeping]
\label{ex:sw-basic-classes-k3-elliptic-blowup}
For a K3 surface,
\[
  \chi=24,\qquad \sigma=-16,\qquad 2\chi+3\sigma=0.
\]
The only Seiberg-Witten basic class is \(K=0\), with invariant \(\pm1\) after
a homology-orientation choice.  Hence the simple-type dimension condition is
saturated exactly:
\[
  K^2=0=2\chi+3\sigma.
\]

For the relatively minimal simply connected elliptic surface \(E(n)\) with
no multiple fibers and \(n\ge2\),
\[
  \chi(E(n))=12n,\qquad \sigma(E(n))=-8n,
  \qquad 2\chi+3\sigma=0.
\]
Let \(F\) be the fiber class.  Since \(F^2=0\), the classes
\begin{equation}
  K_j=(n-2-2j)F,\qquad j=0,\ldots,n-2,
  \label{eq:elliptic-basic-classes}
\end{equation}
all have square zero.  With an orientation convention fixed, their
Seiberg-Witten coefficients are encoded by
\begin{equation}
  \sum_K {\rm SW}_{E(n)}(K)\,\ee^{K}
  =
  \bigl(\ee^F-\ee^{-F}\bigr)^{n-2}.
  \label{eq:elliptic-sw-generating-function}
\end{equation}
The case \(n=2\) is K3.  The formula records both the location of the basic
classes and their binomial coefficients; changing the homology orientation
multiplies all coefficients by a common sign.

If \(\widetilde X=X\#\overline{\mathbb{CP}}^{\,2}\) and \(E\) is the
exceptional class with \(E^2=-1\), then
\[
  \chi(\widetilde X)=\chi(X)+1,
  \qquad
  \sigma(\widetilde X)=\sigma(X)-1.
\]
Thus \(2\chi+3\sigma\) decreases by one.  If \(K^2=2\chi(X)+3\sigma(X)\),
then
\begin{equation}
  (K\pm E)^2
  =
  K^2-1
  =
  2\chi(\widetilde X)+3\sigma(\widetilde X).
  \label{eq:blowup-simple-type-square}
\end{equation}
The blow-up formula for Seiberg-Witten invariants gives
\begin{equation}
  {\rm SW}_{\widetilde X}(K+E)
  =
  {\rm SW}_{\widetilde X}(K-E)
  =
  {\rm SW}_X(K)
  \label{eq:sw-blowup-basic-classes}
\end{equation}
up to the same global orientation convention.  This example is a useful
normalization check: the dimension formula, basic-class positions, and
coefficient doubling are independent pieces of data.
\end{example}

\section{Stable Cohomotopy Refinement And Furuta's Constraint}

The ordinary Seiberg-Witten invariant is an integer obtained after choosing
perturbations and counting or pairing with a finite-dimensional moduli space.
There is a sharper object before this integer is extracted.  After Sobolev
completion and Coulomb-slice gauge fixing, the monopole equations define a
Fredholm map with compact nonlinear part, called the monopole map.  Its
finite-dimensional approximations define a stable cohomotopy class.  The
ordinary Seiberg-Witten invariant is obtained from this stable cohomotopy
class by applying the appropriate Hurewicz-type map and evaluating on the
orientation class.

\paragraph{Monopole map before taking its degree.}
Fix a background connection \(B_0\) on \(L_{\mathfrak s}\), write
\(B=B_0+b\), and impose the Coulomb condition \(d^*b=0\).  After Sobolev
completion, the Seiberg-Witten equations are encoded by a map
\begin{equation}
  \mathcal F:
  L^2_k\!\left(i\Omega^1(X)\oplus\Gamma(S^+_{\mathfrak s})\right)
  \longrightarrow
  L^2_{k-1}\!\left(
    i\Omega^0(X)\oplus i\Omega^{2,+}(X)\oplus
    \Gamma(S^-_{\mathfrak s})
  \right)
  \label{eq:monopole-map-sobolev}
\end{equation}
of the form
\begin{equation}
  \mathcal F(b,q)
  =
  \left(
    d^*b,\,
    d^+b-\sigma(q)+F_{B_0}^+-\eta^+,\,\,
    \not D_{B_0}q+\rho(b)q
  \right).
  \label{eq:monopole-map-components}
\end{equation}
Thus
\begin{equation}
  \mathcal F=L+C ,
  \qquad
  L=d^*\oplus d^+\oplus \not D_{B_0},
  \label{eq:monopole-map-linear-compact-split}
\end{equation}
where \(L\) is Fredholm and \(C\) is compact on bounded sets after Sobolev
embedding and Rellich compactness.  The Weitzenbock estimate in the previous
section supplies the a priori bound needed to restrict to a large ball whose
boundary contains no zeroes of \(\mathcal F\), after the perturbation has been
chosen away from reducible walls.  Choose finite-dimensional subspaces
\[
  W_\Lambda\subset
  L^2_{k-1}\!\left(
    i\Omega^0\oplus i\Omega^{2,+}\oplus\Gamma(S^-_{\mathfrak s})
  \right)
\]
spanned by low eigenmodes of \(LL^*\), and set
\[
  V_\Lambda=L^{-1}(W_\Lambda).
\]
For sufficiently large \(\Lambda\), the projected map
\begin{equation}
  \mathcal F_\Lambda:
  V_\Lambda\longrightarrow W_\Lambda
  \label{eq:finite-dimensional-monopole-map}
\end{equation}
has the same zero set as \(\mathcal F\) in the chosen ball and gives a based
map between one-point compactifications
\begin{equation}
  S^{V_\Lambda}\longrightarrow S^{W_\Lambda}.
  \label{eq:monopole-map-representation-spheres}
\end{equation}
Increasing \(\Lambda\) suspends this map by finite-dimensional pieces of the
Fredholm operator.  The stable homotopy class of the resulting map is the
object refined by Bauer and Furuta.  The ordinary integer invariant is obtained
only after applying a cohomological orientation map to this stable class; it is
therefore a shadow, not the primary object.

\begin{quotedtheorem}[Bauer-Furuta]
\label{qt:bauer-furuta-refinement}
For a closed oriented smooth four-manifold with a fixed
\(\operatorname{Spin}^c\) structure, the finite-dimensional approximation
of the Seiberg-Witten monopole map defines a stable cohomotopy invariant
\begin{equation}
  {\rm BF}_X(\mathfrak s).
  \label{eq:bauer-furuta-class}
\end{equation}
The ordinary Seiberg-Witten invariant is the image of
\({\rm BF}_X(\mathfrak s)\) under the corresponding stable cohomotopy to
cohomology map, with the usual orientation conventions.
\end{quotedtheorem}

The theorem packages the independence of all auxiliary choices in the
construction above: Sobolev index, background connection, radius of the
isolating ball, spectral cutoff, and generic perturbation inside a chamber.
The proof is not a formal consequence of the integer Seiberg-Witten invariant.
It uses properness on the chosen bounded region, homotopy invariance of
finite-dimensional approximations, and compatibility of the suspension
coordinates with the Fredholm index.  The monograph uses this theorem as a
pure four-manifold input adjacent to QFT, while the displayed monopole-map
construction records exactly what additional structure an eventual
non-perturbative twisted-QFT construction would have to recover.

\begin{quotedtheorem}[Furuta \(10/8+2\) inequality]
\label{qt:furuta-ten-eight}
Let \(X\) be a closed smooth spin four-manifold with \(b_2^+(X)>0\).  Then
\begin{equation}
  b_2(X)\ge \frac{10}{8}\,|\sigma(X)|+2.
  \label{eq:furuta-ten-eight}
\end{equation}
\end{quotedtheorem}

\paragraph{Why stable cohomotopy can imply a smooth constraint.}
When \(X\) is spin, the spinor bundles carry a quaternionic structure and the
monopole map is equivariant for the group \(\operatorname{Pin}(2)\subset
\mathbb H^\times\) generated by \(U(1)\) and an element \(j\) with
\(jzj^{-1}=\bar z\).  The finite-dimensional approximation therefore produces
a \(\operatorname{Pin}(2)\)-equivariant stable map between representation
spheres.  The virtual difference of these representation spheres is fixed by
the Dirac index and by \(b_2^+(X)\); for a spin four-manifold the Dirac index
is \(-\sigma(X)/8\) as a quaternionic index.  Furuta's theorem applies
equivariant stable homotopy, in particular the \(\operatorname{Pin}(2)\)
representation-theoretic obstruction to such a stable map, to force
\eqref{eq:furuta-ten-eight}.  Thus the inequality is not derived from a
numerical monopole count.  It uses the extra equivariant information retained
by the Bauer-Furuta refinement.

The relevance for the present chapter is conceptual and structural.  The
Donaldson-Witten path integral suggests that four-manifold invariants should
arise from a cohomological field theory, but ordinary integer invariants can
forget stable homotopy information in the monopole map.  The Bauer-Furuta
class records part of that information.  Furuta's inequality is a theorem
extracted from this stronger structure in the spin case; for K3,
\[
  b_2=22,\qquad |\sigma|=16,
\]
and \eqref{eq:furuta-ten-eight} is saturated:
\[
  22=\frac{10}{8}\cdot16+2.
\]
The monograph treats these as rigorous four-manifold theorems adjacent to
the QFT comparison.  A future theorem-level QFT construction should explain
which refinement of the twisted path integral sees the stable cohomotopy
class rather than only its integer shadow.

\section{The Coulomb-Branch Description}

The Donaldson-Witten theory is the twist of pure \(\mathcal N=2\) Yang--Mills
theory.  Its local supersymmetric Coulomb-branch data for gauge algebra
\(\mathfrak{su}(2)\) were developed in
Chapter~\ref{ch:susy-four-dimensional-n2-seiberg-witten}.  The gauge-invariant
coordinate is
\[
  u=\langle\operatorname{tr}\phi^2\rangle ,
\]
and the special coordinates \((a_D(u),a(u))\) determine the Abelian coupling
\[
  \tau(u)=\frac{da_D/du}{da/du}.
\]
The Coulomb branch has singular fibers where a BPS monopole or dyon becomes
massless.  Near such a fiber, the Wilsonian effective theory contains an
Abelian vector multiplet together with the light charged hypermultiplet.  The
topological twist of this local Abelian theory has localization equations of
Seiberg-Witten monopole type.

The cohomological path integral on a four-manifold receives three kinds of
terms:
\[
\begin{array}{ll}
  \text{instanton description} &
  \text{intersection pairing on ASD moduli spaces},\\[2mm]
  \text{Coulomb branch} &
  \text{integral over }u\text{ with Abelian zero modes and contact terms},\\[2mm]
  \text{singular fibers} &
  \text{monopole or dyon Seiberg-Witten invariants}.
\end{array}
\]
For \(b_2^+(X)>1\), the Coulomb-branch integral has no chamber-wall
contribution in the generic situation used for the simple-type comparison.
For \(b_2^+(X)=1\), the \(u\)-plane contribution and wall crossing are part of
the invariant.  The comparison requires the contact terms of the Abelian
effective action because products of descent observables collide in the
low-energy description.

\section{Chambers, Reducibles, And Wall Crossing}

When \(b_2^+(X)=1\), a metric \(g\) determines a ray
\begin{equation}
  \mathbb R_{>0}\,\omega_g
  \subset
  \{v\in H^2(X;\mathbb R)\mid Q_X(v,v)>0\}
  \label{eq:period-point-ray}
\end{equation}
by the self-dual harmonic line.  A chamber is a connected component of the
positive cone after removing the walls on which reducible connections can
occur.  To describe those walls without hiding the bundle data, take a
rank-two Hermitian bundle \(E\to X\) with fixed determinant and write
\[
  w=c_1(E),\qquad
  p=c_1(E)^2-4c_2(E)=p_1(\mathfrak{su}(E)).
\]
A reducible splitting has the form
\[
  E=L\oplus(\det E\otimes L^{-1}),
\]
and the associated characteristic wall class is
\begin{equation}
  \lambda=2c_1(L)-c_1(E).
  \label{eq:donaldson-wall-class}
\end{equation}
It satisfies
\begin{equation}
  \lambda\equiv w\pmod2,\qquad \lambda^2=p.
  \label{eq:donaldson-wall-class-constraints}
\end{equation}
The reducible connection is anti-self-dual precisely when the self-dual
harmonic projection of \(\lambda\) vanishes, equivalently
\begin{equation}
  \langle \lambda,\omega_g\rangle=0
  \label{eq:donaldson-wall-equation}
\end{equation}
in the \(b_2^+=1\) chamber picture.

\begin{quotedtheorem}[Donaldson wall crossing in link form]
\label{qt:donaldson-wall-crossing-link-form}
Let a generic one-parameter path of metrics cross exactly one wall
\(W_\lambda\) transversely and let \(z\) be a homogeneous Donaldson insertion
whose degree matches the virtual dimension.  The jump of Donaldson
invariants is the oriented link pairing at the reducible stratum:
\begin{equation}
  D_{X,+}^w(z)-D_{X,-}^w(z)
  =
  \varepsilon(\lambda)\,
  \int_{\mathbb L_\lambda}
  \mu_\lambda(z).
  \label{eq:donaldson-wall-crossing-link}
\end{equation}
Here \(\mathbb L_\lambda\) is the link of the reducible in the
parameterized compactified moduli space, \(\mu_\lambda(z)\) is the
restriction of the universal \(\mu\)-class expression to this link, and
\(\varepsilon(\lambda)\) is the orientation sign determined by the direction
in which the period point crosses
\(\langle\lambda,\omega\rangle=0\).  Equivalently, the right hand side is the
Kotschick-Morgan wall-crossing polynomial
\begin{equation}
  {\rm WC}_\lambda^X(z)
  =
  \varepsilon(\lambda)
  \int_{\mathbb L_\lambda}\mu_\lambda(z),
  \label{eq:kotschick-morgan-link-definition}
\end{equation}
whose coefficients are universal functions of the intersection form, the
class \(\lambda\), the lift \(w\), and the degree of \(z\).
\end{quotedtheorem}

\paragraph{Mechanism of the link formula.}
Let
\[
  \mathcal W
  =
  \{(t,A)\mid F_{A,g_t}^+=0\}/\mathcal G
\]
be the parameterized moduli problem for the metric path \(g_t\).  Away from
reducibles and compactification strata, \(\mathcal W\) is the zero set of a
Fredholm section whose index is one larger than the index of the fixed-metric
ASD problem.  Truncate small neighborhoods of the reducible stratum and of the
other Uhlenbeck boundary strata.  The resulting oriented space has boundary
of the form
\begin{equation}
  \partial\mathcal W^{\rm tr}
  =
  \mathcal M_{+}
  \sqcup
  (-\mathcal M_{-})
  \sqcup
  \mathbb L_\lambda
  \sqcup
  \{\text{other compactification faces}\},
  \label{eq:parameterized-wall-crossing-boundary}
\end{equation}
with the last faces absent or cancelled under the hypotheses of the theorem.
The universal Donaldson class representing \(z\) extends over
\(\mathcal W^{\rm tr}\) and is closed there.  Stokes' theorem therefore gives
\[
  0
  =
  \int_{\partial\mathcal W^{\rm tr}}\mu(z)
  =
  D_{X,+}^w(z)-D_{X,-}^w(z)
  +\varepsilon_{\rm link}\int_{\mathbb L_\lambda}\mu_\lambda(z),
\]
which is \eqref{eq:donaldson-wall-crossing-link} after fixing the convention
\(\varepsilon(\lambda)=-\varepsilon_{\rm link}\).  The sign is the orientation
comparison between the outward normal of the deleted tubular neighborhood and
the direction in which the period point crosses
\(\langle\lambda,\omega\rangle=0\).

The local model of the deleted neighborhood is obtained by decomposing the
bundle near a reducible splitting
\[
  E=L\oplus(\det E\otimes L^{-1}).
\]
The diagonal part gives the Abelian stabilizer and the wall equation; the
off-diagonal part gives the normal ASD complex.  Its finite-dimensional
kernel, cokernel, and stabilizer quotient determine the link
\(\mathbb L_\lambda\).  Restricting the universal \(\mu\)-classes to this
normal model leaves only the intersection form, the wall class \(\lambda\), the
lift \(w\), and the topological numbers entering the index.  This is the
geometric reason that the right hand side can be written as the universal
Kotschick-Morgan polynomial rather than as a new invariant attached to each
metric path.

This is the form of the wall-crossing theorem that is most useful for QFT.
It identifies the jump with an actual boundary integral.  In BV language, the
two chamber invariants are the two sides of a one-parameter family of
pushforwards, and the link \(\mathbb L_\lambda\) is the boundary component
where the pushforward fails to be closed.  Thus wall crossing is not an
exception to metric independence; it is the boundary term required by
the Uhlenbeck BV-boundary calculation above.

\begin{remark}[Kotschick-Morgan polynomial normal form]
\label{rem:kotschick-morgan-polynomial-normal-form}
In the situation of
Theorem~\ref{qt:donaldson-wall-crossing-link-form}, restrict for simplicity to
insertions generated by a point class \(x\) and a surface source
\(h\in H_2(X;\mathbb R)\).  For each pair \((r,n)\) with the correct
Donaldson degree, the wall-crossing functional has the universal form
\begin{equation}
  {\rm WC}_\lambda^X(x^r h^n)
  =
  \varepsilon(\lambda)
  \sum_{\substack{j,\ell\ge0\\j+2\ell\le n}}
  c_{r,n;j,\ell}(p,w^2,\lambda^2,\chi,\sigma)\,
  \langle \lambda,h\rangle^{\,n-j-2\ell}
  Q_X(h,h)^\ell
  \langle w,h\rangle^j .
  \label{eq:km-polynomial-normal-form}
\end{equation}
The coefficients are universal rational numbers after the Donaldson
normalization and orientation convention have been fixed.  The allowed
monomials are exactly the scalar polynomials in \(h\) built from the link data
\((Q_X,\lambda,w)\) with total Donaldson surface degree \(2n\).

The link of a reducible is constructed from the normal complex to the
reducible stratum and the universal \(\mu\)-classes restricted to it.  After
the determinant line and orientation of the crossing are fixed, the only
homological tensors available on \(H_2(X;\mathbb R)\) are the intersection
form \(Q_X\), the wall class \(\lambda\), and the fixed lift \(w\).  Therefore
the restriction of a homogeneous insertion \(x^r h^n\) to the link is an
ordinary polynomial in the three scalar contractions
\[
  Q_X(h,h),\qquad \langle\lambda,h\rangle,\qquad \langle w,h\rangle .
\]
The exponent of \(Q_X(h,h)\) consumes two surface insertions, while each
linear contraction consumes one.  This gives the degree constraint
\(j+2\ell\le n\).  Universality follows because the local model of the
reducible link depends on the four-manifold only through the topological
numbers \(p,w^2,\lambda^2,\chi,\sigma\) and the intersection pairing.  The
theorem of Kotschick and Morgan computes these universal coefficients by
evaluating the link integral; \eqref{eq:km-polynomial-normal-form} is the
coordinate form of that computation needed for comparison with the
\(u\)-plane boundary.
\end{remark}

\paragraph{Chamber jump as a BV boundary term.}
Let \(\rho_t\) be a regulated Donaldson-Witten BV semidensity along a generic
metric path \(g_t\), \(t\in[-1,1]\), crossing a single reducible wall
\(\lambda\) at \(t=0\).  Assume the parameterized moduli problem is smooth
away from the reducible link and all insertion classes have been extended to
the truncated compactification.  Then
\begin{equation}
  D_{X,+}^w(z)-D_{X,-}^w(z)
  =
  \mathcal B_{\mathbb L_\lambda}(\rho z),
  \label{eq:wall-crossing-bv-boundary}
\end{equation}
with the sign convention of
\eqref{eq:donaldson-wall-crossing-link}.  If the path crosses several walls,
the jump is the signed sum of the corresponding link functionals.

Under these hypotheses, Stokes' theorem applies to the parameterized
compactified moduli space with small collar neighborhoods of reducibles
removed.  The ordinary boundary at
\(t=1\) gives \(D_{X,+}^w(z)\), the ordinary boundary at \(t=-1\) gives
\(-D_{X,-}^w(z)\), and the new boundary is the link
\(\mathbb L_\lambda\).  In the finite-dimensional BV model, include the
metric parameter among the residual variables and apply
\eqref{eq:singular-stratum-localized-qme-obstruction} to the residual cycle
with the reducible collar removed.  Its boundary functional is precisely the
localized integral over \(\mathbb L_\lambda\).  The link orientation is
exactly the sign \(\varepsilon(\lambda)\) in the wall-crossing theorem.

\section{Observable Map And Contact Terms}

The Donaldson generating function is organized by a point variable \(p\) and
a two-cycle source \(S\in H_2(X;\mathbb R)\):
\[
  Z_D(p,S)
  =
  \left\langle
    \exp\bigl(p\,I(x)+I(S)\bigr)
  \right\rangle .
\]
Here \(I(x)\) has degree four and \(I(S)\) has degree two.  In the Abelian
Coulomb-branch description, the point observable maps to a function of the
Coulomb coordinate,
\[
  I(x)\longmapsto u ,
\]
whereas the surface observable has a linear term in the Abelian field strength
and a quadratic contact term:
\[
  I(S)\longmapsto
  \int_S \Omega_u
  +\frac{1}{2}(S,S)\,T(u).
\]
The two-form \(\Omega_u\) is the appropriate descent two-form in the Abelian
effective theory, and \(T(u)\) is fixed by compatibility of the low-energy
Ward identities, duality transformations, and short-distance collision of two
surface descendants.  The symbol \((S,S)\) denotes the intersection pairing
on \(H_2(X;\mathbb R)\).

At a monopole singularity, the Abelian surface insertion evaluates on the
monopole line bundle and gives an exponential weight
\[
  \exp\bigl((S,\lambda)\bigr)
\]
for a characteristic class \(\lambda=c_1(\mathfrak s)\) associated with the
corresponding \(\operatorname{Spin}^c\) structure, multiplied by the contact
factor coming from \(T(u)\).  The final comparison formula in simple-type
chambers has the general form
\[
  Z_D(p,S)
  =
  e^{Q(S)/2}
  \sum_{\mathfrak s}
  C_{\mathfrak s}\,
  \operatorname{SW}(\mathfrak s)\,
  \exp\bigl(p\,u_{\mathfrak s}+(S,\lambda_{\mathfrak s})\bigr),
\]
where \(Q(S)=(S,S)\).  The constants \(C_{\mathfrak s}\), the phases, and the
choice of chamber depend on the global form of the gauge group, the
orientation convention, and the contact-term normalization.  In this
monograph the displayed expression is treated as a target comparison statement:
the rigorous QFT problem is to derive every term from a constructed twisted
continuum theory and its Wilsonian flow, rather than to declare the formula
from moduli-space geometry alone.

\begin{hypothesis}[Witten simple-type comparison statement]
\label{hyp:witten-simple-type-comparison-datum}
Let \(X\) be a closed connected oriented smooth four-manifold with
\(b_1(X)=0\), \(b_2^+(X)>1\), and Seiberg-Witten simple type.  Fix the lift
\(w\), homology orientation, point-observable convention
\eqref{eq:donaldson-simple-type-relation}, and the trace-delta gauge
normalization used in this chapter.  The Donaldson-Seiberg-Witten comparison
predicts that the reduced Donaldson series has the form
\begin{equation}
  \mathbb D_X^w(h)
  =
  \exp\!\left(\frac{1}{2}Q_X(h,h)\right)
  \sum_{K\in\mathcal B_{\rm SW}(X)}
  A_X^w(K)\,
  {\rm SW}_X(K)\,
  \exp\!\bigl(\langle K,h\rangle\bigr),
  \label{eq:witten-simple-type-comparison}
\end{equation}
where \(\mathcal B_{\rm SW}(X)\) is the set of Seiberg-Witten basic classes.
The phase has the characteristic form
\begin{equation}
  A_X^w(K)
  =
  C_X\,
  (-1)^{\frac{1}{2}(w^2+\langle w,K\rangle)}
  \label{eq:witten-comparison-phase}
\end{equation}
when \(w\) is chosen so that the exponent is integral.  The universal
constant \(C_X\) depends only on the Euler characteristic, signature, and
normalization of Donaldson observables; in the common Moore-Witten
normalization it is written as
\begin{equation}
  C_X=2^{\,2+(7\chi(X)+11\sigma(X))/4}.
  \label{eq:moore-witten-common-constant}
\end{equation}
Equation \eqref{eq:witten-simple-type-comparison} is a theorem in the
mathematical Donaldson/SW comparison under the hypotheses of the relevant
comparison theorems, but in this monograph it is also a QFT target: a
non-perturbative construction of the twisted four-dimensional gauge theory
should derive the same coefficient, phase, contact, and boundary data from
Wilsonian flow.
\end{hypothesis}

The distinction between
\eqref{eq:donaldson-simple-type-exponential-form} and
\eqref{eq:witten-simple-type-comparison} is important.  The first is the
Donaldson simple-type structure: there are some Donaldson basic classes with
some coefficients.  The second identifies those classes and coefficients with
Seiberg-Witten monopole invariants.  The QFT explanation of that
identification is the central RG statement.

\paragraph{Factor ledger for the simple-type comparison.}
Equation~\eqref{eq:witten-simple-type-comparison} contains several pieces
with different origins.  Keeping them separate prevents the comparison from
being read as a single formal substitution:
\begin{enumerate}
\item The Donaldson simple-type theorem supplies only the finite exponential
shape
\[
  \exp(Q_X/2)\sum_K a_K^w\exp\langle K,h\rangle
\]
after the point-operator normalization is fixed.  It does not identify the
labels \(K\) with monopole basic classes or compute the coefficients
\(a_K^w\).
\item The Seiberg-Witten finite-dimensional theory supplies the candidate
labels \(K=c_1(\mathfrak s)\), the oriented counts \({\rm SW}_X(K)\), and the
simple-type vanishing of nonzero-dimensional monopole moduli spaces.  It does
not by itself produce the Donaldson contact Gaussian, the \(w\)-dependent
phase, or the universal power of \(2\).
\item The Abelian low-energy observable map supplies the Gaussian
\(\exp(Q_X(h,h)/2)\) through the surface-contact term \(T(u)\).  This is a
collision-of-descendants datum in the twisted Abelian theory, not a
consequence of the finite monopole count.
\item Electromagnetic-duality determinant lines and the lift \(w\) supply the
phase
\((-1)^{(w^2+\langle w,K\rangle)/2}\).  This phase must be matched with the
orientation conventions for Donaldson and monopole determinant lines; changing
the lift changes the exceptional-class parity as in
Example~\ref{ex:donaldson-side-k3-elliptic-blowup-tables}.
\item The numerical constant \(C_X\) comes from the regularized Abelian
measure factors, one-loop gravitational terms, and the normalization of
Donaldson observables.  Its dependence on \(\chi(X)\) and \(\sigma(X)\) is
therefore a determinant-line and contact-normalization claim, not a
topological fact following from simple type.
\item Chamber dependence is controlled separately.  For \(b_2^+>1\) the
simple-type comparison uses the absence of a residual regular \(u\)-plane
bulk contribution under the stated hypotheses.  For \(b_2^+=1\), the
regularized \(u\)-plane boundary and the Donaldson link wall-crossing formula
must be matched before a chamber-specific comparison is meaningful.
\end{enumerate}
The open problem below is precisely to construct a single QFT framework in
which all six entries arise from the same regulator, anomaly line,
Wilsonian flow, and compactification boundary convention.

\paragraph{Proof obligations hidden by the comparison formula.}
A derivation of Hypothesis~\ref{hyp:witten-simple-type-comparison-datum} from
QFT would have to prove the following chain of constructions and
identifications.  First, one needs a non-perturbative construction of the
twisted pure \(\mathcal N=2\) gauge theory on \(X\), including a regulator or
BV/lattice replacement for the gauge-fixed measure, the scalar charge \(Q\),
the descent observables \(I(x)\) and \(I(S)\), and compactification boundary
data for the ASD localization locus.  Second, the Wilsonian flow must be
defined on this cohomological theory in a way compatible with \(Q\) and with
the anomaly line, and must produce the Seiberg-Witten Coulomb-branch family
away from singular fibers with the Abelian contact term \(T(u)\).  Third, the
regular \(u\)-plane integral has to be shown to supply only the stated
chamber-dependent boundary contribution, with no hidden bulk term for the
\(b_2^+>1\) simple-type situation.  Fourth, a neighborhood of each singular
fiber must be identified with the twisted Abelian monopole theory, including
orientation lines, determinant phases, and the map from surface observables to
the monopole determinant line.  This is the step that turns the singular-fiber
boundary contribution into \({\rm SW}_X(K)\exp\langle K,h\rangle\).  Finally,
the normalization of the point observable, the Gaussian surface contact term,
the phase
\((-1)^{(w^2+\langle w,K\rangle)/2}\), and the power of \(2\) in
\eqref{eq:moore-witten-common-constant} must be derived from the same
regularized determinant and contact-term conventions, rather than fitted after
the fact.

Thus \eqref{eq:witten-simple-type-comparison} is not used below as an
unproved physical theorem.  It is a precise target statement: each factor in the
formula names a separate analytic, topological, or Wilsonian ingredient that a
future theorem-level construction must supply.

\begin{example}[Donaldson-side tables from the comparison statement]
\label{ex:donaldson-side-k3-elliptic-blowup-tables}
The following table is not an independent derivation of the
Donaldson--Seiberg-Witten comparison.  It is the explicit Donaldson-side output
of Hypothesis~\ref{hyp:witten-simple-type-comparison-datum} in examples where
the Seiberg-Witten basic classes were fixed in
Example~\ref{ex:sw-basic-classes-k3-elliptic-blowup}.  We take \(w=0\) when
this is compatible with the displayed parity and choose the common
Moore--Witten constant \eqref{eq:moore-witten-common-constant}.
\[
\begin{array}{c|c|c}
  X & \sum_K{\rm SW}_X(K)e^K & \mathbb D_X^0(h)\\
  \hline
  K3
  & 1
  & \exp(Q_X(h,h)/2)\\[1mm]
  E(n)
  & (e^F-e^{-F})^{n-2}
  & 2^{2-n}\exp(Q_X(h,h)/2)(e^{\langle F,h\rangle}
       -e^{-\langle F,h\rangle})^{n-2}
\end{array}
\]
For \(E(2)=K3\), the second row reduces to the first.  For \(E(3)\), the
Donaldson basic classes predicted by the comparison are \(\pm F\) with
opposite coefficients in this \(w=0\) convention:
\begin{equation}
  \mathbb D_{E(3)}^0(h)
  =
  \frac12\,e^{Q(h,h)/2}
  \bigl(e^{\langle F,h\rangle}-e^{-\langle F,h\rangle}\bigr).
  \label{eq:donaldson-e3-table}
\end{equation}
If \(\widetilde X=X\#\overline{\mathbb{CP}}^{\,2}\), let \(E\) be the
exceptional cohomology class and choose \(e\in H_2(\widetilde X;\mathbb R)\)
with \(Q_{\widetilde X}(e,e)=-1\) and \(\langle E,e\rangle=1\).  With the
lift chosen so that the two exceptional basic classes have the same phase,
the comparison gives
\begin{equation}
  \mathbb D_{\widetilde X}^{\widetilde w}(h+te)
  =
  e^{-t^2/2}\cosh(t)\,\mathbb D_X^w(h).
  \label{eq:donaldson-blowup-cosh-form}
\end{equation}
With the adjacent lift convention the two exceptional phases differ and the
factor \(\cosh(t)\) is replaced by \(\sinh(t)\), up to the common orientation
sign.  In the exponential generating series this gives an explicit parity
table:
\begin{align}
  [t^{2m}]\,e^{-t^2/2}\cosh t
  &=
  \sum_{r=0}^{m}
  \frac{(-1)^r}{2^r r!(2m-2r)!},
  &
  [t^{2m+1}]\,e^{-t^2/2}\cosh t
  &=0,
  \label{eq:donaldson-blowup-even-lift-coefficients}\\
  [t^{2m+1}]\,e^{-t^2/2}\sinh t
  &=
  \sum_{r=0}^{m}
  \frac{(-1)^r}{2^r r!(2m+1-2r)!},
  &
  [t^{2m}]\,e^{-t^2/2}\sinh t
  &=0.
  \label{eq:donaldson-blowup-odd-lift-coefficients}
\end{align}
Thus the even lift has coefficients \(1,0,-1/12,\ldots\) in degrees
\(0,2,4,\ldots\), while the adjacent lift has coefficients \(1,-1/3,1/20,\ldots\)
in degrees \(1,3,5,\ldots\).  These numbers are not new invariants; they are
the exact exceptional-insertion Taylor coefficients obtained by multiplying
the contact Gaussian by the two exceptional basic-class phases.  The point of
writing the table is that the constants, phases, Gaussian factor, basic-class
locations, and blow-up square shift are separate pieces of information; none
is determined by the word ``simple type'' alone.
\end{example}

\section{The \texorpdfstring{\(u\)-Plane}{u-Plane} Integral And Boundary Contributions}

For \(b_2^+(X)=1\), the Coulomb-branch contribution cannot be suppressed by
the simple-type argument.  The low-energy path integral over the regular part
of the \(u\)-plane is written, after fixing the Donaldson observable
normalization used in \eqref{eq:witten-simple-type-comparison}, as
\begin{equation}
  Z_u^w(p,S;\omega)
  =
  \int_{\mathcal B_u^{\rm reg}}
  da\wedge d\bar a\,
  A(u)^{\chi(X)}B(u)^{\sigma(X)}
  \Psi_{\Gamma,w,\omega}(\tau,\bar\tau;S)
  \exp\!\left(2pu+S^2T(u)\right).
  \label{eq:u-plane-integral-data}
\end{equation}
Here \(\mathcal B_u^{\rm reg}\) is the Coulomb branch with the singular
fibers removed, \(a\) is a local special coordinate,
\[
  \tau(u)=\frac{da_D/du}{da/du},
\]
\(\Gamma=H^2(X;\mathbb Z)/{\rm torsion}\) with intersection pairing \(Q_X\),
and \(\omega\) is the chamber period point in
\eqref{eq:period-point-ray}.  The functions
\begin{equation}
  A(u)=\alpha\left(\frac{du}{da}\right)^{1/2},
  \qquad
  B(u)=\beta\,\Delta(u)^{1/8}
  \label{eq:u-plane-measure-factors}
\end{equation}
encode the gravitational measure factors of the Abelian twisted effective
theory; \(\Delta(u)\) is the discriminant of the Seiberg-Witten family and
\(\alpha,\beta\) are convention-dependent constants.  The factor
\(\Psi_{\Gamma,w,\omega}\) is the Siegel-Narain theta kernel obtained by
summing over Abelian fluxes shifted by \(w/2\), with the positive and
negative projections defined by \(\omega\).  Its dependence on \(\omega\) is
the analytic source of wall crossing.  The contact term \(T(u)\) is the same
surface-observable collision term that appears in the observable map above.

\begin{definition}[Regularized \(u\)-plane integral]
\label{def:regularized-u-plane-integral}
Let \(\Omega_u^w(p,S;\omega)\) denote the two-form integrand in
\eqref{eq:u-plane-integral-data}.  Choose small disjoint disks
\[
  D_\epsilon(u_\alpha)
\]
around the singular fibers \(u_\alpha\) of the Seiberg-Witten family and, in a
coordinate \(q=e^{2\pi i\tau}\) near the weak-coupling cusp, remove the region
\(|q|<\epsilon_\infty\).  The regulated domain is
\[
  \mathcal B_{u,\epsilon}^{\rm reg}
  =
  \mathcal B_u
  \setminus
  \left(
    \bigcup_\alpha D_\epsilon(u_\alpha)
    \cup\{|q|<\epsilon_\infty\}
  \right).
\]
The regulated Coulomb contribution is
\begin{equation}
  Z_{u,\epsilon}^w(p,S;\omega)
  =
  \int_{\mathcal B_{u,\epsilon}^{\rm reg}}
  \Omega_u^w(p,S;\omega).
  \label{eq:regulated-u-plane-integral}
\end{equation}
A renormalized \(u\)-plane contribution means a limit
\begin{equation}
  Z_u^w
  =
  \lim_{\epsilon\to0}
  \left[
    Z_{u,\epsilon}^w
    +\sum_\alpha C_{\alpha,\epsilon}^{\rm sing}
    +C_{\infty,\epsilon}^{\rm cusp}
  \right],
  \label{eq:renormalized-u-plane-integral}
\end{equation}
when the boundary counterterms \(C_{\alpha,\epsilon}^{\rm sing}\) and
\(C_{\infty,\epsilon}^{\rm cusp}\) have been derived from the corresponding
monopole/dyon and weak-coupling effective theories.  They are not arbitrary
subtractions; they are part of the declared low-energy regularization datum.
\end{definition}

\paragraph{Wall-normal factor in the theta kernel.}
Only one elementary analytic mechanism is needed to understand why the
\(u\)-plane wall crossing is localized on fluxes satisfying the wall
equation.  Fix a lattice flux
\[
  \lambda\in\Gamma+\frac{w}{2}
\]
and write
\[
  x_\lambda(\omega)=\langle\lambda,\omega\rangle
\]
for its component along the positive period ray.  The complete
Siegel-Narain kernel contains smooth factors in \(u,S\), and a wall-normal
factor whose local model is
\begin{equation}
  E_t(x)=\operatorname{erf}(\sqrt{\pi t}\,x),
  \qquad t>0.
  \label{eq:u-plane-wall-normal-erf}
\end{equation}
The parameter \(t\) is proportional to the positive imaginary part of the
Abelian coupling in the normal direction.  This model records the only
feature relevant for the wall jump: as \(t\to\infty\),
\[
  E_t(x)\to \operatorname{sgn}(x)
  \qquad(x\ne0),
\]
while
\begin{equation}
  \frac{dE_t}{dx}
  =
  2\sqrt t\,\exp(-\pi t x^2)
  \label{eq:u-plane-wall-normal-delta-sequence}
\end{equation}
is a delta sequence of total mass \(2\):
\[
  \int_{-\infty}^{\infty}2\sqrt t\,\exp(-\pi t x^2)\,\dd x=2.
\]
Thus changing the chamber by moving \(\omega\) through
\(x_\lambda(\omega)=0\) changes the corresponding completed sign factor by
exactly \(2\).  Flux sectors with \(x_\lambda(\omega)\ne0\) remain smooth in
\(\omega\); the only singular chamber dependence comes from the codimension
one locus \(x_\lambda=0\).  This is the local analytic content behind the
phrase that wall crossing is supported on reducible fluxes.  The global
statement still requires the remaining factors in the \(u\)-plane integrand,
the singular-fiber counterterms, and the determinant-line orientation
matching to be controlled.

\begin{controlledapproximation}[Wall crossing from the \(u\)-plane boundary]
\label{ca:u-plane-wall-crossing-boundary}
Assume that the Abelian low-energy description is valid on
\(\mathcal B_u^{\rm reg}\), that the singular-fiber contributions are
computed by the corresponding monopole or dyon Seiberg-Witten theories, and
that the theta kernel has the standard Gaussian decay away from walls.  When
\(\omega\) crosses the wall \(\langle\lambda,\omega\rangle=0\), the
discontinuity of \(Z_u^w\) is a boundary term supported by the flux sector
\(\lambda\), and equals the link wall-crossing functional
{\rm WC}\({}_\lambda^X\) after the contact normalization and determinant-line
phase are matched.

Only the positive/negative decomposition of the lattice vector in
\(\Psi_{\Gamma,w,\omega}\) changes discontinuously at a wall.  The local
model \eqref{eq:u-plane-wall-normal-erf}--\eqref{eq:u-plane-wall-normal-delta-sequence}
shows that the chamber jump of a completed sign factor is a delta-supported
normal contribution with total jump \(2\).  Away from \(\lambda_+=0\), the
Gaussian factor in the theta kernel gives a smooth
\(\omega\)-dependent integral.  Crossing the wall changes the prescription
for the sector with \(\lambda_+=0\); differentiating with respect to the
period point produces a total \(\bar\partial\)-term on the \(u\)-plane.  By
Stokes' theorem the change is therefore a sum of boundary contributions from
the cusps and singular fibers.  The singular-fiber boundaries are precisely
the monopole/dyon effective theories, while the cusp and reducible boundary
combine into the Donaldson wall link.  Matching determinant-line orientations
and the contact term \(T(u)\) identifies the result with
\eqref{eq:kotschick-morgan-link-definition}.  This proves the conditional
statement: the analytic equality rests on the stated validity of the Abelian
Wilsonian description and on the matching of its measure factors.
\end{controlledapproximation}

\section{Comparison As A QFT Statement}

The comparison can be formulated as a precise datum.  Fix:
\[
\begin{array}{ll}
  \text{UV theory} &
  \text{twisted pure }\mathcal N=2\text{ }SU(2)\text{ gauge theory on }X,\\[1mm]
  \text{observable algebra} &
  Q\text{-cohomology generated by Donaldson descent classes},\\[1mm]
  \text{metric chamber} &
  \text{a chamber for }b_2^+(X)=1\text{ or the }b_2^+>1\text{ regime},\\[1mm]
  \text{Wilsonian coordinate} &
  u\text{ and the Abelian special geometry near each singular fiber},\\[1mm]
  \text{contact prescription} &
  T(u)\text{ and all collision counterterms for descent insertions},\\[1mm]
  \text{IR localization} &
  \operatorname{Spin}^c\text{ monopole moduli spaces with orientations}.
\end{array}
\]
The QFT comparison asserts equality between the ultraviolet cohomological
correlators and the sum of the Coulomb-branch integral plus monopole
contributions computed from this datum.  A theorem-level derivation would
therefore contain the following ingredients:
\[
\begin{array}{ll}
  \text{construction} &
  \text{a four-dimensional twisted continuum gauge theory},\\[1mm]
  \text{Ward identities} &
  Q\text{-invariance, descent, and metric variation including contacts},\\[1mm]
  \text{RG control} &
  \text{flow from the UV description to the Abelian Coulomb chart},\\[1mm]
  \text{singular fiber analysis} &
  \text{derivation of the monopole effective theory and its measure},\\[1mm]
  \text{compactness} &
  \text{finite-dimensional compactified moduli spaces and boundary terms},\\[1mm]
  \text{orientation and phases} &
  \text{matching of determinant lines and electromagnetic duality phases}.
\end{array}
\]
The finite-dimensional Donaldson and Seiberg-Witten invariants are rigorous
objects once their compactifications, orientations, perturbations, and wall
crossing conventions are fixed.  The physical comparison supplies a deep
explanation by RG, and the present formulation isolates the additional
analytic QFT construction needed to turn that explanation into a theorem
inside four-dimensional quantum field theory.

\begin{openproblem}[Donaldson-SW QFT comparison]
\label{op:donaldson-seiberg-witten-qft-comparison}
Construct the Donaldson-Witten theory and its Coulomb-branch Wilsonian flow
with enough analytic control to derive the Seiberg-Witten comparison formula,
including the ultraviolet twisted gauge theory, the Abelian monopole
effective theories at singular fibers, the \(u\)-plane contribution, descent
contact terms, determinant-line phases, chamber dependence, wall crossing,
and compactification boundary contributions.
\end{openproblem}
