\chapter{BF Theory}
\label{ch:tqft-bf-theory}

\ConventionsEuclidean

BF theory is the first gauge-theoretic example in this volume.  It already
displays recurring ingredients of topological gauge theory:
fields with form degree, ordinary gauge symmetry, higher-form gauge symmetry,
reducibility, metric-free classical action, extended observables, and a BV
complex.  The abelian compact theory also has a finite cochain model, so some
correlation functions reduce to exact finite sums.

\section{Nonabelian Classical Fields}

Let \(M\) be a closed oriented smooth \(D\)-manifold.  Let \(G\) be a compact
Lie group, let \(P\to M\) be a principal \(G\)-bundle, and let
\[
  \operatorname{ad}P=P\times_G\mathfrak g
\]
be the adjoint bundle.  Choose an invariant nondegenerate pairing
\[
  \langle-,-\rangle:\mathfrak g^*\otimes\mathfrak g\longrightarrow\mathbb R .
\]
The classical fields of \(D\)-dimensional nonabelian BF theory are a
connection \(A\) on \(P\) and a \(B\)-field
\[
  B\in\Omega^{D-2}(M,\operatorname{ad}P^*).
\]
The curvature of \(A\) is
\[
  F_A=\dd A+\frac12[A,A]\in\Omega^2(M,\operatorname{ad}P),
\]
where the bracket combines the Lie bracket on \(\mathfrak g\) with exterior
wedge product.  The action is
\begin{equation}
  S_{\rm BF}[A,B]
  =
  2\pi\ii\int_M \langle B\wedge F_A\rangle .
  \label{eq:bf-action-nonabelian}
\end{equation}
The coefficient is fixed so that the compact abelian specialization below has
integer periods in the exponent.

\begin{proposition}[Euler--Lagrange equations]
\label{prop:bf-euler-lagrange}
On a closed \(M\), the stationary equations of
\eqref{eq:bf-action-nonabelian} are
\[
  F_A=0,
  \qquad
  \dd_A B=0 .
\]
\end{proposition}

\begin{proof}
Let \(\alpha\in\Omega^1(M,\operatorname{ad}P)\) and
\(\beta\in\Omega^{D-2}(M,\operatorname{ad}P^*)\) be variations of \(A\) and
\(B\).  The curvature variation is
\[
  \left.\frac{\dd}{\dd t}\right|_{t=0}F_{A+t\alpha}
  =
  \dd_A\alpha .
\]
Therefore
\[
  \delta S_{\rm BF}
  =
  2\pi\ii\int_M
  \left(
    \langle\beta\wedge F_A\rangle
    +
    \langle B\wedge\dd_A\alpha\rangle
  \right).
\]
The first term gives \(F_A=0\) because \(\beta\) is arbitrary.  For the second
term, use the covariant Leibniz rule
\[
  \dd\langle B\wedge\alpha\rangle
  =
  \langle\dd_A B\wedge\alpha\rangle
  +
  (-1)^{D-2}\langle B\wedge\dd_A\alpha\rangle .
\]
The integral of the left side vanishes on a closed manifold, so
\[
  \int_M\langle B\wedge\dd_A\alpha\rangle
  =
  (-1)^{D-1}\int_M\langle\dd_A B\wedge\alpha\rangle .
\]
Since \(\alpha\) is arbitrary, \(\dd_A B=0\).
\end{proof}

\section{Gauge Symmetry And Reducibility}

The ordinary gauge group \(\mathcal G(P)\) acts by
\[
  A\longmapsto A^g=g^{-1}Ag+g^{-1}\dd g,
  \qquad
  B\longmapsto B^g=\operatorname{Ad}^*_{g^{-1}}B,
\]
where the second formula is the coadjoint action; after choosing an invariant
inner product on \(\mathfrak g\) it is written as \(B^g=g^{-1}Bg\).
Infinitesimally, for \(c\in\Omega^0(M,\operatorname{ad}P)\),
\[
  \delta_c A=\dd_Ac,
  \qquad
  \delta_c B=[B,c].
\]
There is also a \(B\)-field shift symmetry.  For
\[
  \eta\in\Omega^{D-3}(M,\operatorname{ad}P^*)
\]
define
\[
  \delta_\eta A=0,
  \qquad
  \delta_\eta B=\dd_A\eta .
\]

\begin{proposition}[Invariance of the shift symmetry]
\label{prop:bf-shift-invariance}
On a closed manifold,
\[
  S_{\rm BF}[A,B+\dd_A\eta]-S_{\rm BF}[A,B]
  =
  2\pi\ii\int_M \dd\langle\eta\wedge F_A\rangle .
\]
Hence the shift is a symmetry on closed \(M\).  On a manifold with boundary it
produces the boundary term
\[
  2\pi\ii\int_{\partial M}\langle\eta\wedge F_A\rangle .
\]
\end{proposition}

\begin{proof}
The change in the action is
\[
  2\pi\ii\int_M\langle\dd_A\eta\wedge F_A\rangle .
\]
The covariant Leibniz rule and Bianchi identity \(\dd_A F_A=0\) give
\[
  \dd\langle\eta\wedge F_A\rangle
  =
  \langle\dd_A\eta\wedge F_A\rangle
  +(-1)^{D-3}\langle\eta\wedge\dd_A F_A\rangle
  =
  \langle\dd_A\eta\wedge F_A\rangle .
\]
This proves the formula.
\end{proof}

The shift symmetry is reducible.  If
\(\eta=\dd_A\eta_1\), then
\[
  \delta_\eta B=\dd_A^2\eta_1=F_A\cdot\eta_1 .
\]
Thus the transformation vanishes on the flat-connection locus.  Continuing,
\[
  \eta_1\in\Omega^{D-4}(M,\operatorname{ad}P^*),
  \quad
  \eta_1\sim\eta_1+\dd_A\eta_2,
  \quad
  \ldots ,
\]
ending in degree \(0\).  In the abelian theory \(\dd_A=\dd\) and the
reducibility holds off shell because \(\dd^2=0\).  The BV field space
therefore contains the ordinary ghost \(c\), the \(B\)-shift ghost
\(\eta\), the ghosts-for-ghosts \(\eta_1,\eta_2,\ldots\), and antifields for
all these variables.  This is the finite-stage reducible gauge system whose
BV integration is modeled by Chapter~\ref{ch:tqft-bv-integration-localization}.

\section{Compact Abelian Normalization}

A precise continuum normalization of compact abelian BF theory uses
differential characters.  For a \(U(1)\) one-form gauge field \(A\), the
holonomy around a closed one-cycle \(\gamma\) is
\[
  W_q(\gamma)=
  \exp\!\left(2\pi\ii q\int_\gamma A\right),
  \qquad q\in\mathbb Z .
\]
For a compact \((D-2)\)-form gauge field \(B\), the surface operator on a
closed \((D-2)\)-cycle \(\Sigma\) is
\[
  V_p(\Sigma)=
  \exp\!\left(2\pi\ii p\int_\Sigma B\right),
  \qquad p\in\mathbb Z .
\]
The local action is
\[
  S_{\rm BF}[A,B]=2\pi\ii N\int_M B\wedge\dd A ,
  \qquad N\in\mathbb Z_{>0}.
\]
Large gauge transformations shift the periods of \(A\) and \(B\) by
integers, and the factor \(N\) makes the exponent single-valued.  The
equation obtained by varying \(B\) is
\[
  N\,\dd A=0
\]
as a differential character.  Thus the theory describes a flat
\(\mathbb Z_N\) one-form gauge sector after the continuous modes and gauge
redundancies are accounted for.

\section{Finite \texorpdfstring{\(\mathbb Z_N\)}{ZN} Cochain Model}

Let \(K\) be a finite cell decomposition of a closed connected oriented
\(D\)-manifold \(M\).  Write \(C^r=C^r(K;\mathbb Z_N)\), and let
\[
  \delta_r:C^r\longrightarrow C^{r+1}
\]
be the cellular coboundary.  The finite constraint form of \(\mathbb Z_N\)
BF theory has a one-cochain gauge field \(a\in C^1\) and a Lagrange
multiplier \(b\in\operatorname{Hom}(C^2,\mathbb Z_N)\).  Its normalized
finite sum is
\begin{equation}
  Z_{\rm BF}^{(N)}(K)
  =
  \frac{1}{|C^0|\,|C^2|}
  \sum_{a\in C^1}\sum_{b\in\operatorname{Hom}(C^2,\mathbb Z_N)}
  \exp\!\left(\frac{2\pi\ii}{N}\,b(\delta_1a)\right).
  \label{eq:finite-bf-cochain-partition}
\end{equation}
The factor \(|C^0|^{-1}\) divides by ordinary \(0\)-form gauge
transformations \(a\mapsto a+\delta_0\lambda\).  The factor \(|C^2|^{-1}\)
normalizes the \(b\)-sum as a finite Fourier transform.

\begin{proposition}[Finite BF partition function]
\label{prop:finite-bf-partition-function}
For a connected \(K\),
\[
  Z_{\rm BF}^{(N)}(K)
  =
  \frac{|Z^1(K;\mathbb Z_N)|}{|C^0(K;\mathbb Z_N)|}
  =
  \frac{|H^1(K;\mathbb Z_N)|}{|H^0(K;\mathbb Z_N)|}
  =
  \frac{|H^1(M;\mathbb Z_N)|}{N}.
\]
\end{proposition}

\begin{proof}
The finite character identity
\[
  \frac{1}{|C^2|}
  \sum_{b\in\operatorname{Hom}(C^2,\mathbb Z_N)}
  \exp\!\left(\frac{2\pi\ii}{N}\,b(x)\right)
  =
  \begin{cases}
  1,&x=0,\\
  0,&x\ne0
  \end{cases}
\]
holds for every \(x\in C^2\).  Applying it to \(x=\delta_1a\) reduces
\eqref{eq:finite-bf-cochain-partition} to
\[
  Z_{\rm BF}^{(N)}(K)=|C^0|^{-1}\,|\ker\delta_1|.
\]
Since
\[
  H^1(K;\mathbb Z_N)=\ker\delta_1/\operatorname{im}\delta_0,
\]
we have \(|\ker\delta_1|=|H^1|\,|\operatorname{im}\delta_0|\).  Also
\[
  H^0(K;\mathbb Z_N)=\ker\delta_0
\]
and \(|C^0|=|\ker\delta_0|\,|\operatorname{im}\delta_0|\).  Hence
\[
  |C^0|^{-1}|\ker\delta_1|
  =
  |H^1|/|H^0|.
\]
For connected \(K\), \(H^0(K;\mathbb Z_N)\cong\mathbb Z_N\), giving the last
formula.
\end{proof}

The value \(|H^1|/N\) is the groupoid cardinality of flat
\(\mathbb Z_N\)-bundles on a connected manifold: each flat bundle has
automorphism group \(\mathbb Z_N\), and groupoid cardinality weights each
isomorphism class by the reciprocal of the automorphism-group order.

\section{Wilson Operators And Linking}

Let \(\gamma\in Z_1(K;\mathbb Z)\) be a closed cellular one-cycle.  Its
Wilson operator of charge \(q\in\mathbb Z_N\) is
\[
  W_q(\gamma)(a)
  =
  \exp\!\left(\frac{2\pi\ii q}{N}\langle a,\gamma\rangle\right).
\]
Gauge invariance follows from the cellular Stokes formula:
\[
  \langle a+\delta_0\lambda,\gamma\rangle
  =
  \langle a,\gamma\rangle+\langle\lambda,\partial\gamma\rangle
  =
  \langle a,\gamma\rangle .
\]

Let \(\Sigma\) be a closed oriented \((D-2)\)-cycle, represented on the dual
cellulation by a two-cochain \(\sigma_\Sigma\in C^2(K;\mathbb Z_N)\) through
Poincare duality.  The \(B\)-surface insertion of charge \(p\in\mathbb Z_N\)
shifts the finite Fourier constraint by
\[
  b(\delta a)
  \longmapsto
  b(\delta a+p\,\sigma_\Sigma).
\]
Thus the \(b\)-sum imposes
\[
  \delta a=-p\,\sigma_\Sigma .
\]
This equation is the finite cochain form of a prescribed surface singularity.

\begin{proposition}[Linking phase in finite BF theory]
\label{prop:finite-bf-linking-phase}
Assume \(M=S^D\) and let \(\gamma\) and \(\Sigma\) be disjoint closed cycles
of dimensions \(1\) and \(D-2\).  Choose a two-chain \(C\) with
\(\partial C=\gamma\), and define
\[
  \operatorname{Lk}(\gamma,\Sigma)
  =
  \langle\sigma_\Sigma,C\rangle\in\mathbb Z .
\]
With the sign convention above,
\[
  \langle W_q(\gamma)V_p(\Sigma)\rangle
  =
  \exp\!\left(
    -\frac{2\pi\ii pq}{N}\operatorname{Lk}(\gamma,\Sigma)
  \right)
  \langle V_p(\Sigma)\rangle .
\]
\end{proposition}

\begin{proof}
After inserting \(V_p(\Sigma)\), the \(b\)-sum enforces
\[
  \delta a=-p\,\sigma_\Sigma .
\]
For any solution \(a\),
\[
  \langle a,\gamma\rangle
  =
  \langle a,\partial C\rangle
  =
  \langle\delta a,C\rangle
  =
  -p\,\langle\sigma_\Sigma,C\rangle .
\]
Therefore the Wilson insertion contributes the phase
\[
  \exp\!\left(
    -\frac{2\pi\ii pq}{N}\operatorname{Lk}(\gamma,\Sigma)
  \right)
\]
for every allowed \(a\).  The remaining sum is the sum defining
\(\langle V_p(\Sigma)\rangle\).  Independence of the choice of \(C\) follows
because \(H_2(S^D;\mathbb Z)=0\) for \(D\ne2\), and in the relevant
codimension-linking case \(D\ge3\).
\end{proof}

\section{Metric Independence And Boundary Data}

The classical action \eqref{eq:bf-action-nonabelian} uses the orientation of
\(M\), the exterior derivative, wedge product, bundle data, and invariant
pairing.  It uses no Riemannian metric.  Quantum metric independence requires
additional data:
\[
  \text{BV field space},\qquad
  \text{gauge-fixing Lagrangian},\qquad
  \text{regularized semidensity},\qquad
  \text{anomaly analysis}.
\]
If the metric variation of the gauge-fixed integrand is BV-exact and if the
BV Stokes boundary term vanishes, then the corresponding BV-closed
correlators are metric independent by
Corollary~\ref{cor:bv-gauge-fixing-independence}.  On a manifold with
boundary, Proposition~\ref{prop:bf-shift-invariance} shows that the shift
symmetry produces boundary phase-space data.  The boundary fields are the
restrictions of \(A\) and \(B\), and the boundary symplectic pairing is
obtained from the boundary term in the variation:
\[
  \Omega_{\partial M}
  =
  2\pi\ii\int_{\partial M}\langle\delta B\wedge\delta A\rangle .
\]
A functorial BF theory must choose a boundary polarization or boundary
condition, then prove compatibility with gluing.

\begin{openproblem}[Continuum nonabelian BF theory with boundaries]
\label{op:continuum-bf-boundaries}
Construct nonabelian continuum BF theory on manifolds with boundary as a
functorial BV-BFV theory.  The construction must specify the boundary phase
space, residual fields, gauge fixing, anomaly line, gluing pairing, and the
relationship between Wilson, surface, and boundary observables.
\end{openproblem}
