\chapter{BV Integration And Finite-Dimensional Localization}
\label{ch:tqft-bv-integration-localization}

\ConventionsEuclidean

Topological and cohomological field theories use two related operations.  The
first is BV integration: a semidensity on an odd symplectic space is restricted
to a Lagrangian cycle and integrated.  The second is cohomological
localization: an integral of a closed density is deformed by an exact term and
evaluated from the zero locus of the deformation.  This chapter proves the
finite-dimensional identities with all cycle, boundary, and determinant data
visible.  Infinite-dimensional gauge-theoretic applications in later chapters
must supply a regulator, a replacement for compactness, determinant-line
orientations, and control of singular strata.

\section{Odd Symplectic BV Data}

Let \(\mathcal F\) be a finite-dimensional smooth graded manifold.  Its
fermionic coordinates are coordinates of the structure sheaf; they are not
points of an underlying ordinary set.  An odd symplectic form of cohomological
degree \(-1\) is a closed nondegenerate two-form
\[
  \omega\in\Omega^2(\mathcal F),
  \qquad
  |\omega|=-1 .
\]
For a homogeneous function \(F\in C^\infty(\mathcal F)\), the Hamiltonian
vector field \(X_F\) is defined by
\[
  \iota_{X_F}\omega=\delta F .
\]
The BV bracket is the odd Poisson bracket
\[
  \{F,G\}=X_F(G).
\]
It has cohomological degree \(1\), satisfies the graded Jacobi identity, and
obeys the graded Leibniz rule in each argument.  If \(S\) is an even function
of degree zero, the vector field
\[
  Q_S=\{S,-\}
\]
has degree \(1\).  The classical master equation is
\begin{equation}
  \{S,S\}=0 .
  \label{eq:bv-cme-finite-dimensional}
\end{equation}
The Jacobi identity gives
\[
  Q_S^2F=\frac12\{\{S,S\},F\},
\]
so \eqref{eq:bv-cme-finite-dimensional} is equivalent to \(Q_S^2=0\) on
functions.

In local Darboux coordinates \((x^i,\xi_i)\), where
\[
  |x^i|+|\xi_i|=1,\qquad
  \omega=\sum_i \delta x^i\,\delta\xi_i,
\]
the bracket is the coordinate expression determined by the preceding
Hamiltonian convention.  A choice of Berezinian density \(\mu\) compatible
with these coordinates gives a BV Laplacian on functions,
\begin{equation}
  \Delta_\mu
  =
  \sum_i(-1)^{|x^i|}
  \frac{\partial^2}{\partial x^i\,\partial\xi_i}.
  \label{eq:bv-laplacian-darboux}
\end{equation}
Intrinsic BV integration is most cleanly stated for semidensities.  In a
coordinate chart a semidensity is written \(f\,\mu^{1/2}\); the canonical
odd-symplectic operator on semidensities is denoted \(\Delta\).  After the
choice of \(\mu\), it is represented by \(\Delta_\mu\) on the coefficient
function \(f\).  The price of working with functions is that a change of
\(\mu\) changes the displayed operator.

\section{Quantum Master Equation And Observables}

Let \(\rho\) be a semidensity on \(\mathcal F\).  The finite-dimensional
quantum master equation is
\begin{equation}
  \Delta\rho=0 .
  \label{eq:finite-bv-qme-semidensity}
\end{equation}
If \(\rho=\exp(S/\hbar)\mu^{1/2}\) in a chosen compatible Berezinian
trivialization, then \eqref{eq:finite-bv-qme-semidensity} becomes
\begin{equation}
  \frac12\{S,S\}+\hbar\Delta_\mu S=0 .
  \label{eq:finite-bv-qme-action}
\end{equation}
Indeed, applying the second-order operator \(\Delta_\mu\) to
\(\exp(S/\hbar)\) gives
\[
  \Delta_\mu\!\left(\ee^{S/\hbar}\right)
  =
  \hbar^{-1}
  \left(\Delta_\mu S+\frac{1}{2\hbar}\{S,S\}\right)\ee^{S/\hbar},
\]
which is equivalent to \eqref{eq:finite-bv-qme-action}.

An observable is a function \(\mathcal O\) for which
\[
  \Delta(\mathcal O\rho)=0 .
\]
In the same trivialization, if \(S\) obeys
\eqref{eq:finite-bv-qme-action}, this condition reads
\begin{equation}
  \Delta_\mu\mathcal O+\hbar^{-1}\{S,\mathcal O\}=0 .
  \label{eq:bv-quantum-observable}
\end{equation}
The classical limit of \eqref{eq:bv-quantum-observable} is
\(\{S,\mathcal O\}=0\), the \(Q_S\)-closedness condition.

\begin{remark}[BV product identity]
\label{rem:bv-product-identity}
For a homogeneous function \(F\) and an action \(S\) satisfying
\eqref{eq:finite-bv-qme-action},
\[
  \Delta_\mu(F\ee^{S/\hbar})
  =
  \left(\Delta_\mu F+\hbar^{-1}\{S,F\}\right)\ee^{S/\hbar}.
\]

The defining second-order identity for \(\Delta_\mu\) is
\[
  \Delta_\mu(FG)
  =
  (\Delta_\mu F)G
  +(-1)^{|F|}F\Delta_\mu G
  +(-1)^{|F|}\{F,G\}.
\]
Take \(G=\ee^{S/\hbar}\).  Since \(S\) is even,
\[
  \{F,\ee^{S/\hbar}\}
  =
  \hbar^{-1}\{F,S\}\ee^{S/\hbar}.
\]
The graded antisymmetry of the odd bracket converts the last term to
\(\hbar^{-1}\{S,F\}\ee^{S/\hbar}\) with the displayed convention.  The term
containing \(\Delta_\mu G\) vanishes by the quantum master equation for
\(\ee^{S/\hbar}\mu^{1/2}\).  This proves the identity.
\end{remark}

\section{Lagrangian Cycles And BV Stokes}

A Lagrangian submanifold \(\mathcal L\subset\mathcal F\) is a graded
submanifold for which the pullback of \(\omega\) vanishes and whose dimension
is half the dimension of \(\mathcal F\) in the graded sense.  The restriction
of a semidensity \(\rho\) to \(\mathcal L\) is an ordinary Berezinian density
on \(\mathcal L\), denoted \(\rho|_{\mathcal L}\).  The BV expectation of an
observable is
\[
  \langle\mathcal O\rangle_{\mathcal L}
  =
  \int_{\mathcal L}(\mathcal O\rho)|_{\mathcal L}.
\]
The Lagrangian \(\mathcal L\) is the gauge-fixing cycle.

\begin{theorem}[Finite-dimensional BV Stokes theorem]
\label{thm:finite-bv-stokes}
Let \(\sigma\) be a compactly supported semidensity on an odd symplectic
finite-dimensional graded manifold \(\mathcal F\), and let
\(\mathcal L\subset\mathcal F\) be an oriented Lagrangian submanifold without
boundary.  Then
\[
  \int_{\mathcal L}(\Delta\sigma)|_{\mathcal L}=0 .
\]
If \(\mathcal L\) has boundary, the same calculation gives the induced
boundary integral on \(\partial\mathcal L\).
\end{theorem}

\begin{proof}
Work in Darboux coordinates for which
\(\mathcal L=\{\xi_i=0\}\) on a coordinate patch.  Write
\[
  \sigma=f(x,\xi)\,\mu^{1/2}.
\]
The restriction of \(\Delta\sigma\) to \(\mathcal L\) is represented by
\[
  \left.
  \sum_i(-1)^{|x^i|}
  \frac{\partial^2 f}{\partial x^i\,\partial\xi_i}
  \right|_{\xi=0}
  \dd x .
\]
This is the ordinary divergence on \(\mathcal L\) of the vector density with
components
\[
  (-1)^{|x^i|}
  \left.\frac{\partial f}{\partial\xi_i}\right|_{\xi=0}.
\]
Its integral over a boundaryless compact patch is zero by Stokes' theorem,
and the patchwise statements glue because the canonical operator on
semidensities is coordinate invariant.  For a Lagrangian with boundary the
ordinary Stokes theorem gives exactly the boundary term.
\end{proof}

\begin{remark}[Gauge-fixing independence]
\label{rem:bv-gauge-fixing-independence}
Let \(\rho\) satisfy \(\Delta\rho=0\), and let
\(\mathcal O\) satisfy \(\Delta(\mathcal O\rho)=0\).  Let
\(\mathcal L_s\), \(s\in[0,1]\), be a smooth family of compact Lagrangian
cycles such that the swept chain has no boundary except
\(\mathcal L_1-\mathcal L_0\).  Then
\[
  \int_{\mathcal L_1}(\mathcal O\rho)|_{\mathcal L_1}
  =
  \int_{\mathcal L_0}(\mathcal O\rho)|_{\mathcal L_0}.
\]
The family \(\mathcal L_s\) sweeps a Lagrangian cobordism, so this is the
direct finite-dimensional consequence of
Theorem~\ref{thm:finite-bv-stokes} applied to the closed semidensity
\(\mathcal O\rho\).  The hypotheses \(\Delta\rho=0\) and
\(\Delta(\mathcal O\rho)=0\) are the finite-dimensional BV analogues of the
quantum master equation and the observable Ward identity; they are exactly
what makes the integrand closed for the Stokes argument.
\end{remark}

\section{BV Pushforward}

Let the odd symplectic space split as
\[
  \mathcal F=\mathcal F_{\rm res}\times\mathcal F_{\rm fluc},
  \qquad
  \omega=\omega_{\rm res}+\omega_{\rm fluc}.
\]
The first factor contains residual fields and the second factor contains
fluctuation fields to be integrated out.  Let
\(\mathcal L_{\rm fluc}\subset\mathcal F_{\rm fluc}\) be a Lagrangian cycle.
For a semidensity \(\rho\) on \(\mathcal F\), define the BV pushforward
\[
  (\pi_*\rho)(\phi_{\rm res})
  =
  \int_{\mathcal L_{\rm fluc}}
  \rho(\phi_{\rm res},\phi_{\rm fluc})\big|_{\mathcal L_{\rm fluc}} .
\]
The result is a semidensity on \(\mathcal F_{\rm res}\).

\begin{theorem}[QME preservation by BV pushforward]
\label{thm:bv-pushforward-qme-preservation}
Assume that \(\rho\) has compact support in the fluctuation directions, or
that the chosen fluctuation cycle has decay strong enough to eliminate the
boundary term in Theorem~\ref{thm:finite-bv-stokes}.  If
\[
  \Delta_{\mathcal F}\rho=0,
\]
then
\[
  \Delta_{\rm res}(\pi_*\rho)=0 .
\]
The same statement holds for observables:
\[
  \Delta_{\mathcal F}(\mathcal O\rho)=0
  \quad\Longrightarrow\quad
  \Delta_{\rm res}\pi_*(\mathcal O\rho)=0 .
\]
\end{theorem}

\begin{proof}
For the product odd symplectic form,
\[
  \Delta_{\mathcal F}=\Delta_{\rm res}+\Delta_{\rm fluc}.
\]
Therefore
\[
  \Delta_{\rm res}\pi_*\rho
  =
  \int_{\mathcal L_{\rm fluc}}
  (\Delta_{\rm res}\rho)|_{\mathcal L_{\rm fluc}}
  =
  -\int_{\mathcal L_{\rm fluc}}
  (\Delta_{\rm fluc}\rho)|_{\mathcal L_{\rm fluc}} .
\]
The last integral is zero by Theorem~\ref{thm:finite-bv-stokes} applied in
the fluctuation variables.  If \(\Delta_{\mathcal F}(\mathcal O\rho)=0\),
the identical decomposition gives
\[
  \Delta_{\rm res}\pi_*(\mathcal O\rho)
  =
  -\int_{\mathcal L_{\rm fluc}}
  (\Delta_{\rm fluc}(\mathcal O\rho))|_{\mathcal L_{\rm fluc}},
\]
and the fluctuation BV Stokes theorem again makes the right hand side vanish.
\end{proof}

\begin{definition}[Boundary functional of a fluctuation cycle]
\label{def:bv-fluctuation-boundary-functional}
Let \(\mathcal L_{\rm fluc}\) be a fluctuation Lagrangian cycle with boundary.
For a semidensity \(\sigma=f\,\mu^{1/2}\) on
\(\mathcal F_{\rm fluc}\), define
\(\mathcal B_{\partial\mathcal L_{\rm fluc}}(\sigma)\) by
\[
  \int_{\mathcal L_{\rm fluc}}(\Delta_{\rm fluc}\sigma)|_{\mathcal L_{\rm fluc}}
  =
  \mathcal B_{\partial\mathcal L_{\rm fluc}}(\sigma).
  \label{eq:bv-boundary-functional-definition}
\]
In a Darboux chart in which
\(\mathcal L_{\rm fluc}=\{\xi_i=0\}\), this boundary functional is the
ordinary Stokes boundary integral of the vector density with components
\[
  (-1)^{|x^i|}
  \left.
  \frac{\partial f}{\partial \xi_i}
  \right|_{\xi=0}.
  \label{eq:bv-boundary-vector-density}
\]
The coordinate-independence is precisely the boundary version of
Theorem~\ref{thm:finite-bv-stokes}.
\end{definition}

\begin{proposition}[BV pushforward with fluctuation boundary]
\label{prop:bv-pushforward-boundary-obstruction}
Let \(\mathcal F=\mathcal F_{\rm res}\times\mathcal F_{\rm fluc}\) and let
\(\mathcal L_{\rm fluc}\subset\mathcal F_{\rm fluc}\) be a fluctuation
Lagrangian cycle with boundary.  Assume that all integrations converge and
that \(\rho\) obeys the full finite-dimensional QME
\(\Delta_{\mathcal F}\rho=0\).  Then the pushed semidensity satisfies
\[
  \Delta_{\rm res}(\pi_*\rho)
  =
  -\mathcal B_{\partial\mathcal L_{\rm fluc}}(\rho).
  \label{eq:bv-pushforward-boundary-obstruction}
\]
Thus BV pushforward preserves the QME if and only if the displayed boundary
functional vanishes or is cancelled by an additional boundary, residue, or
regulator contribution.  The same formula holds with \(\rho\) replaced by
\(\mathcal O\rho\) for quantum observables.
\end{proposition}

\begin{proof}
As in Theorem~\ref{thm:bv-pushforward-qme-preservation},
\(\Delta_{\mathcal F}=\Delta_{\rm res}+\Delta_{\rm fluc}\).  Therefore
\[
  \Delta_{\rm res}\pi_*\rho
  =
  \int_{\mathcal L_{\rm fluc}}
  (\Delta_{\rm res}\rho)|_{\mathcal L_{\rm fluc}}
  =
  -\int_{\mathcal L_{\rm fluc}}
  (\Delta_{\rm fluc}\rho)|_{\mathcal L_{\rm fluc}} .
  \label{eq:bv-boundary-pushforward-proof}
\]
Definition~\ref{def:bv-fluctuation-boundary-functional} identifies the last
integral with the boundary functional.  This proves
\eqref{eq:bv-pushforward-boundary-obstruction}.  The observable statement is
identical after replacing \(\rho\) by \(\mathcal O\rho\).
\end{proof}

In field theory this theorem is the finite-dimensional model for Wilsonian
integration of BV fields.  A regulator must provide a finite-dimensional
replacement or a trace-class analogue of the split above.  If
\(\Delta\rho=\mathcal A\rho\), the semidensity \(\mathcal A\rho\) is the
finite-dimensional anomaly insertion; the pushforward preserves the anomaly
class rather than producing a solution of the quantum master equation.

\section{Singular Strata As BV Boundary Data}
\label{sec:bv-singular-strata-boundary-data}

Gauge-theoretic localization often integrates over a smooth moduli stratum
whose natural compactification has additional strata.  Let
\(\mathcal M^{\rm sm}\) denote the smooth locus and let
\[
  \overline{\mathcal M}
  =
  \mathcal M^{\rm sm}\sqcup\bigsqcup_\alpha S_\alpha
  \label{eq:moduli-compactification-strata}
\]
be a stratified compactification.  In applications \(S_\alpha\) may be a
reducible-connection stratum, a singular-monopole stratum, or the
small-instanton stratum of an Uhlenbeck compactification.  A BV localization
argument over \(\mathcal M^{\rm sm}\) is not complete until it states the
boundary functional produced by the strata \(S_\alpha\).

\begin{definition}[Boundary-cancellation data]
\label{def:boundary-cancellation-data}
For a regulated BV localization problem with smooth fixed locus
\(\mathcal M^{\rm sm}\) and compactification
\eqref{eq:moduli-compactification-strata}, boundary-cancellation data
consists of one of the following:
\begin{enumerate}
  \item vanishing of every stratum functional,
  \(\mathcal B_{S_\alpha}(\rho)=0\);
  \item a relative integration cycle whose boundary cancels
  \(\sum_\alpha\mathcal B_{S_\alpha}(\rho)\);
  \item a resolved regulated fixed locus
  \(\widetilde{\mathcal M}\to\overline{\mathcal M}\) with an induced
  semidensity \(\widetilde\rho\);
  \item a finite-dimensional contour or residue rule whose Stokes boundary
  equals \(-\sum_\alpha\mathcal B_{S_\alpha}(\rho)\).
\end{enumerate}
In the resolution case the necessary equality is the identity of pushed
residual semidensities
\[
  \pi_*^{\widetilde{\mathcal M}}\widetilde\rho
  =
  \pi_*^{\mathcal M^{\rm sm}}\rho
  +
  \sum_\alpha C_\alpha,
  \qquad
  \Delta_{\rm res} C_\alpha
  =
  \mathcal B_{S_\alpha}(\rho),
  \label{eq:resolved-boundary-cancellation-data}
\]
where \(C_\alpha\) is the contribution supported over the exceptional or
relative piece replacing \(S_\alpha\).  Equation
\eqref{eq:resolved-boundary-cancellation-data} is data to be proved from
the regulator; it is not implied by writing a compactification symbol.
\end{definition}

\begin{proposition}[Singular-stratum obstruction to localized QME]
\label{prop:singular-stratum-obstruction-localized-qme}
Assume that a regulated gauge-theoretic BV integral has been localized to a
smooth finite-dimensional fixed locus \(\mathcal M^{\rm sm}\) and that the
full regulated semidensity obeys the QME before the limiting
compactification is taken.  If the compactification of
\(\mathcal M^{\rm sm}\) contains strata \(S_\alpha\), then the residual
localized semidensity obeys
\[
  \Delta_{\rm res}(\pi_*^{\mathcal M^{\rm sm}}\rho)
  =
  -\sum_\alpha \mathcal B_{S_\alpha}(\rho).
  \label{eq:singular-stratum-localized-qme-obstruction}
\]
Therefore the localized expression is a BV pushforward satisfying the QME
only after boundary-cancellation data in the sense of
Definition~\ref{def:boundary-cancellation-data} have been supplied and
verified.
\end{proposition}

\begin{proof}
Choose a collar neighborhood of each singular stratum in the compactified
moduli problem and remove a tubular neighborhood of radius \(\epsilon\).
The resulting truncated fixed locus
\(\mathcal M_\epsilon^{\rm sm}\) is a manifold with boundary.  Applying
Proposition~\ref{prop:bv-pushforward-boundary-obstruction} gives
\[
  \Delta_{\rm res}(\pi_*^{\mathcal M_\epsilon^{\rm sm}}\rho)
  =
  -\sum_\alpha
  \mathcal B_{\partial_\alpha\mathcal M_\epsilon^{\rm sm}}(\rho).
  \label{eq:singular-stratum-truncated-boundary}
\]
If the limit \(\epsilon\to0\) exists as a distributional or semidensity
limit, the boundary functionals converge to the stratum functionals
\(\mathcal B_{S_\alpha}(\rho)\), giving
\eqref{eq:singular-stratum-localized-qme-obstruction}.  If the limit does
not exist, the localization formula itself is not yet defined.  A vanishing,
relative-cycle, resolution, or residue input is exactly the additional input
that makes the right hand side vanish or cancel in the residual QME.
\end{proof}

For the \(S^4\) instanton factors of
Chapter~\ref{ch:susy-localization-compact-manifolds}, the Uhlenbeck
small-instanton strata are precisely such \(S_\alpha\).  Replacing the
Uhlenbeck compactification by the Gieseker/Nekrasov resolution supplies a
candidate resolution data.  The statement that the original four-dimensional
path integral selects those resolved data is therefore a regulator theorem or
a hypothesis, as stated in
Hypothesis~\ref{hyp:field-theoretic-gieseker-matching}, rather than a
consequence of formal \(Q\)-exact deformation alone.

\section{Cohomological Localization}

Let \(M\) be an oriented finite-dimensional manifold and let
\(\mathcal A\) be a finite-dimensional graded-commutative algebra.  An
integration functional
\[
  \int_M:\Omega^\bullet(M)\otimes\mathcal A\longrightarrow\mathbb C
\]
is obtained by taking the top ordinary form degree and the chosen Berezin
coefficient in \(\mathcal A\).  Let \(Q\) be an odd derivation of
\(\Omega^\bullet(M)\otimes\mathcal A\) such that \(Q^2=0\).  Assume that the
cycle and boundary conditions give
\begin{equation}
  \int_M Q\beta=0
  \label{eq:localization-q-stokes}
\end{equation}
for every integrand \(\beta\) in the class being integrated.

Let \(S\) and \(\mathcal O\) be even \(Q\)-closed elements and let \(V\) be
odd.  Define
\[
  I(t)=\int_M \mathcal O\,\ee^{-S-tQV}.
\]

\begin{theorem}[\(Q\)-exact deformation invariance]
\label{thm:q-exact-localization-invariance}
Under the hypotheses above,
\[
  \frac{dI}{dt}=0 .
\]
\end{theorem}

\begin{proof}
Differentiate under the integral:
\[
  \frac{dI}{dt}
  =
  -\int_M \mathcal O\,(QV)\,\ee^{-S-tQV}.
\]
Since \(Q\mathcal O=QS=Q^2V=0\), the graded Leibniz rule gives
\[
  Q\!\left(V\mathcal O\,\ee^{-S-tQV}\right)
  =
  (QV)\mathcal O\,\ee^{-S-tQV}.
\]
Thus \(dI/dt\) is minus the integral of a \(Q\)-exact term, and
\eqref{eq:localization-q-stokes} gives the result.
\end{proof}

The theorem has content only after \eqref{eq:localization-q-stokes} has been
verified.  For compact \(M\) and \(Q=\dd\), it is Stokes' theorem.  For
noncompact cycles it is a decay estimate.  For contour prescriptions it is a
statement about the chosen relative homology class.

\section{One-Loop Normal Form}

Assume the real part of \(QV\) is nonnegative and its zero locus is a compact
submanifold \(F\subset M\).  Let \(N_F\) be the normal bundle of \(F\).  In a
tubular neighborhood write \(y\) for even normal coordinates and
\(\chi\) for their odd partners.  Suppose the quadratic normal part is
\begin{equation}
  QV
  =
  \frac12\,y^T A y+\frac12\,\chi^T B\chi+O(3),
  \label{eq:localization-normal-quadratic}
\end{equation}
where \(A\) is a positive definite symmetric operator on the even normal
bundle and \(B\) is a nondegenerate antisymmetric operator on the odd normal
bundle.  The orientation of the normal integration cycle fixes the branch of
\(\sqrt{\det A}\) and the sign of \(\operatorname{Pf}(B)\).

\begin{proposition}[Normal Gaussian factor]
\label{prop:normal-gaussian-localization-factor}
If the limit \(t\to\infty\) may be exchanged with the integral after the
standard rescaling \(y=t^{-1/2}u\), then the normal variables contribute
\[
  \frac{\operatorname{Pf}(B)}{\sqrt{\det A}} .
\]
Consequently the localized integral has the finite-dimensional form
\[
  I
  =
  \int_F
  \frac{\iota_F^*(\mathcal O\,\ee^{-S})}
  {e_Q(N_F)},
\]
where \(e_Q(N_F)\) is the equivariant Euler class represented by the normal
complex.
\end{proposition}

\begin{proof}
The even normal integral is
\[
  \int_{\mathbb R^r}
  \exp\!\left(-\frac{t}{2}y^TAy\right)\dd^r y
  =
  (2\pi)^{r/2}t^{-r/2}(\det A)^{-1/2}.
\]
The odd normal integral is
\[
  \int \dd\chi_1\cdots\dd\chi_{2s}\,
  \exp\!\left(-\frac{t}{2}\chi^TB\chi\right)
  =
  t^s\operatorname{Pf}(B)
\]
with the Berezin orientation fixed by the ordering of the \(\chi\)'s.  In a
cohomological normal complex the odd rank equals the even rank after pairing
by \(Q\), so the powers of \(t\) cancel once the ordinary normalization of
the Thom density is included.  The remaining determinant is
\(\operatorname{Pf}(B)/\sqrt{\det A}\).  The higher-order terms in
\eqref{eq:localization-normal-quadratic} vanish in the rescaled limit under
the stated dominated-convergence hypothesis.  Gluing these local normal
factors over \(F\) gives the inverse equivariant Euler class of \(N_F\).
\end{proof}

\section{Mathai--Quillen Rank-One Model}

The following model fixes the normalization used in localization.  Let
\[
  s_a:\mathbb R\longrightarrow\mathbb R,\qquad s_a(x)=a x,
  \qquad a\ne0 ,
\]
be a section of the oriented trivial real line bundle over \(\mathbb R\).
For \(t>0\), define the one-form
\[
  \Theta_{t,a}
  =
  \left(\frac{t}{2\pi}\right)^{1/2}
  a\,\exp\!\left(-\frac{t a^2x^2}{2}\right)\dd x .
\]
It is the rank-one Mathai--Quillen Thom density pulled back by \(s_a\).  Its
integral is
\begin{equation}
  \int_{\mathbb R}\Theta_{t,a}
  =
  \operatorname{sign}(a).
  \label{eq:rank-one-mq-integral}
\end{equation}
Indeed, setting \(u=\sqrt t\,|a|x\) gives
\[
  \left(\frac{t}{2\pi}\right)^{1/2}
  a\,\dd x
  =
  \operatorname{sign}(a)\frac{\dd u}{\sqrt{2\pi}},
\]
and
\[
  \int_{\mathbb R}\frac{1}{\sqrt{2\pi}}\ee^{-u^2/2}\dd u=1.
\]
This proves \eqref{eq:rank-one-mq-integral}.  The sign is the local degree of
the section at its zero \(x=0\), namely the orientation sign of the derivative
\(\dd s_a=a\,\dd x\).

For a rank-\(r\) oriented Euclidean vector bundle \(E\to M\) with connection
\(\nabla\), section \(s\), and curvature \(R_\nabla\), the Mathai--Quillen
representative is obtained by replacing the single odd variable above by odd
fiber variables \(\chi^a\) and using the exponent
\[
  -\frac{t}{2}\|s\|^2
  +\ii\sqrt t\,\chi_a(\nabla s)^a
  +\frac12\,\chi_a(R_\nabla)^{ab}\chi_b .
\]
The Berezin integral over the \(\chi\)'s gives a closed differential form
whose cohomology class is independent of \(t\).  When the zero locus of \(s\)
is transverse, the \(t\to\infty\) limit is the signed integration current on
\(s^{-1}(0)\).  The rank-one calculation above is the local normal
calculation at a transverse zero.

\section{Atiyah--Bott Fixed-Point Model on the Two-Sphere}
\label{sec:atiyah-bott-s2-model}

The preceding normal determinant has a visible fixed-point form in the
two-sphere model.  Write \(S^2\) with coordinates
\[
  x\in[-1,1],\qquad \varphi\in\mathbb R/2\pi\mathbb Z
\]
away from the two poles, and orient it by
\[
  \omega=\dd\varphi\wedge\dd x .
\]
Then \(\int_{S^2}\omega=4\pi\).  Let the circle act by
\(\varphi\mapsto\varphi+\alpha\), and let \(V=\partial_\varphi\) be its
generating vector field.  With equivariant parameter \(\lambda\), define
\[
  \dd_\lambda:=\dd-\lambda\,\iota_V,
  \qquad
  H=x .
\]
Since \(\iota_V\omega=\dd x\), the equivariant two-form
\[
  \Omega_\lambda:=\omega+\lambda H
\]
is closed:
\[
  \dd_\lambda\Omega_\lambda
  =
  \lambda\,\dd x-\lambda\,\iota_V\omega=0 .
\]

\begin{proposition}[Fixed-point formula in the \(S^2\) model]
\label{prop:atiyah-bott-s2-model}
For \(\lambda\ne0\),
\begin{equation}
  \int_{S^2}\ee^{\Omega_\lambda}
  =
  2\pi\,\frac{\ee^\lambda-\ee^{-\lambda}}{\lambda}.
  \label{eq:s2-fixed-point-integral}
\end{equation}
The same value is obtained from the two fixed points of the circle action:
\[
  2\pi
  \left(
    \frac{\ee^{\lambda H(N)}}{\lambda w_N}
    +
    \frac{\ee^{\lambda H(S)}}{\lambda w_S}
  \right),
  \qquad
  H(N)=1,\quad H(S)=-1,\quad w_N=1,\quad w_S=-1 .
\]
\end{proposition}

\begin{proof}
Only the top ordinary form degree contributes to the integral:
\[
  \ee^{\Omega_\lambda}
  =
  \ee^{\lambda x}(1+\omega),
  \qquad
  [\ee^{\Omega_\lambda}]_{\rm top}
  =
  \ee^{\lambda x}\omega .
\]
Therefore
\[
  \int_{S^2}\ee^{\Omega_\lambda}
  =
  \int_0^{2\pi}\dd\varphi
  \int_{-1}^{1}\ee^{\lambda x}\dd x
  =
  2\pi\,\frac{\ee^\lambda-\ee^{-\lambda}}{\lambda}.
\]

Near the north pole, use a complex coordinate \(z\) in which the circle acts
by \(z\mapsto\ee^{\ii\alpha}z\).  The tangent weight is \(w_N=1\), and
\(H(N)=1\).  Near the south pole the coordinate \(z^{-1}\) has tangent weight
\(w_S=-1\), and \(H(S)=-1\).  The one-dimensional normal Gaussian factor is
the inverse equivariant Euler factor \((\lambda w_p)^{-1}\) at each fixed
point \(p\), multiplied by the angular normalization \(2\pi\).  Summing the
two contributions gives exactly \eqref{eq:s2-fixed-point-integral}.
\end{proof}

This model is finite-dimensional and compact, so no boundary term is hidden.
In gauge-theory localization the fixed set is replaced by a moduli space of
solutions to field equations, the weights become determinant lines of
elliptic complexes, and compactness or a relative-cycle prescription must be
proved before the fixed-point formula has the status of a field-theoretic
identity.

\section{Boundary And Noncompact Terms}

If the integration cycle has boundary or noncompact ends, the derivative in
Theorem~\ref{thm:q-exact-localization-invariance} is the corresponding
boundary integral:
\[
  \frac{dI}{dt}
  =
  -\int_{\partial M_{\rm eff}}
  \text{the boundary value of }V\mathcal O\,\ee^{-S-tQV}.
\]
A localization statement must therefore state one of the following pieces of
data:
\[
\begin{gathered}
  \text{a compact integration cycle},\qquad
  \text{a decay estimate at every noncompact end},\\
  \text{a boundary condition cancelling the term},\qquad
  \text{a specified relative cycle or residue prescription}.
\end{gathered}
\]
In gauge-theoretic localization, Coulomb-branch ends, reducible connections,
small instantons, singular monopole strata, and boundaries of moduli spaces
are exactly the loci where this analysis enters.  A residue prescription such
as a Jeffrey--Kirwan rule is part of the integration-cycle data; it is a
finite-dimensional replacement for a boundary analysis, and its validity in a
field-theoretic construction must be proved from the chosen regulator and
cycle.

\begin{openproblem}[Infinite-dimensional BV localization]
\label{op:infinite-dimensional-bv-localization}
Give theorem-level localization statements for cohomological gauge theories
from a regulated BV-BFV functional integral, including integration cycles,
compactness or boundary estimates, zero modes, determinant lines, singular
strata, and gluing compatibility.
\end{openproblem}
