\chapter{Twists of Supersymmetric Theories}
\label{ch:tqft-twists-of-supersymmetric-theories}

\ConventionsEuclidean

A topological twist is a change of the rotation background of a
supersymmetric QFT by an \(R\)-symmetry background.  Its first output is
representation-theoretic: fields and supercharges are reinterpreted as
sections of different geometric bundles.  Its second output, when the
twisted theory has a scalar odd symmetry \(Q\) satisfying the hypotheses of
Chapter~\ref{ch:tqft-cohomological-field-theories}, is a cohomological QFT.
The two steps are logically distinct.  The twist supplies a candidate
cohomological differential; the Ward identity, stress-tensor relation,
contact prescription, and regulator decide whether it defines a
metric-independent continuum theory.

\section{Twisting as a Background Construction}

Let the flat-space Euclidean theory have rotation group
\(\operatorname{Spin}(d)\) and compact internal \(R\)-symmetry group \(G_R\).
The odd supersymmetry generators form a finite-dimensional
\(\mathbb Z_2\)-odd representation
\[
  \mathsf Q
  \quad\text{of}\quad
  \operatorname{Spin}(d)\times G_R .
\]
The fields form representations of the same group together with gauge and
flavor representations.

\begin{definition}[Twisting homomorphism]
A twisting homomorphism is a continuous homomorphism
\[
  \rho:\operatorname{Spin}(d)\longrightarrow G_R .
\]
The corresponding twisted rotation group is the diagonal subgroup
\[
  \operatorname{Spin}(d)_\rho
  =
  \{(s,\rho(s))\mid s\in\operatorname{Spin}(d)\}
  \subset \operatorname{Spin}(d)\times G_R .
\]
Every flat-space representation \(V\) of
\(\operatorname{Spin}(d)\times G_R\) becomes a representation
\[
  V_\rho
  :=
  {\rm Res}^{\operatorname{Spin}(d)\times G_R}_{\operatorname{Spin}(d)_\rho}V
\]
of the twisted rotation group.
\end{definition}

On a curved oriented Riemannian \(d\)-manifold \(M\), this restriction has the
following geometric meaning.  Let \(P_{\operatorname{Spin}}(M)\) be a spin
frame bundle when \(M\) is spin.  Turning on the \(R\)-symmetry background
determined by the twist means choosing the principal \(G_R\)-bundle
\[
  P_R(M)=P_{\operatorname{Spin}}(M)\times_{\rho}G_R .
\]
If \(V\) is a representation of \(\operatorname{Spin}(d)\times G_R\), the
twisted associated bundle is
\[
  E_V^\rho(M)
  =
  P_{\operatorname{Spin}}(M)\times_{\operatorname{Spin}(d)} V_\rho .
\]
Thus a field which was a spinor in the untwisted flat theory can become a
differential form, a tensor, or a section of another natural bundle after
restriction to \(\operatorname{Spin}(d)_\rho\).

\begin{definition}[Scalar twisting supercharge]
A supercharge \(Q\in\mathsf Q\) is a scalar for the twist \(\rho\) if
\[
  Q\in \mathsf Q^{\operatorname{Spin}(d)_\rho}.
\]
It is a cohomological candidate if, in the twisted algebra of fields or
observables,
\begin{equation}
  Q^2
  =
  \delta_{\rm gauge}(\lambda_Q)+
  \delta_{\rm flavor}(\alpha_Q)+
  \mathcal L_{v_Q}
  \label{eq:twist-q-square-general}
\end{equation}
with specified field-dependent or background-dependent parameters.  On a
chosen observable sector, \(Q\) is a differential exactly when the right side
acts trivially there.
\end{definition}

For a topological twist on a closed manifold, the intended sector in the
examples below is gauge-invariant, flavor-invariant, and invariant under the
bosonic symmetry generated by \(v_Q\).  In the Donaldson-Witten twist below,
\(v_Q=0\) and
\(Q^2\) is a gauge transformation by the adjoint scalar.  In many
localization problems on curved manifolds, \(v_Q\) is an isometry; the theory
is then equivariant with respect to that isometry rather than purely
topological.

\section{Anomaly and Global-Form Conditions}

The homomorphism \(\rho\) is a representation-theoretic input.  A QFT twist
requires additional conditions.

\begin{definition}[Regulator-level twist datum]
A regulator-level twist datum on \(M\) consists of:
\[
\begin{array}{ll}
  \text{rotation geometry} & P_{\operatorname{Spin}}(M)
      \text{ or the corresponding spin}^{c}\text{/oriented replacement},\\
  \text{\(R\)-background} & P_R(M)
      \text{ induced by }\rho\text{ or by a specified refinement},\\
  \text{field/BV space} & \mathcal F_\Lambda^\rho,\\
  \text{odd symmetry} & Q_\Lambda,\\
  \text{closure relation} & Q_\Lambda^2
      \text{ as in \eqref{eq:twist-q-square-general}},\\
  \text{regulated action or state} & S_\Lambda^\rho\text{ or }\omega_\Lambda^\rho,\\
  \text{integration cycle} & \Gamma_\Lambda^\rho,\\
  \text{observable algebra} & \mathcal A_\Lambda^\rho
      \text{ including contact representatives}.
\end{array}
\]
The datum is admissible when the \(R\)-symmetry used in the twist is
nonanomalous for the chosen backgrounds and when \(Q_\Lambda\) preserves
\(S_\Lambda^\rho\), \(\Gamma_\Lambda^\rho\), and the source/contact
prescription.
\end{definition}

The global condition may be weaker than the original spin condition.  For
example, after twisting, every field in a particular multiplet may transform
as a tensor field on an oriented four-manifold.  This tensorial appearance is
a conclusion about the associated bundles after the twist; possible
anomaly-line and global-form constraints still come from fermion
determinants, gauge bundles, discrete theta angles, and background
\(R\)-symmetry bundles.  Each theory requires its own determinant-line and
anomaly check.

\section{Metric Independence from a Twisted Supercharge}

Let \(M\) be compact and let
\[
  Z_g^\rho(\mathcal O_1,\ldots,\mathcal O_n)
\]
be a regulated correlator of \(Q_\Lambda\)-closed insertions in the twisted
theory.  The general cohomological criterion was proved in
Chapter~\ref{ch:tqft-cohomological-field-theories}.  In a supersymmetric
twist, the required metric-response statement is supplied by the
supercurrent multiplet together with the chosen curved-background coupling.

\begin{remark}[Regulator-level metric-independence criterion]
\label{rem:twist-metric-independence}
Assume the regulator-level twist datum above satisfies:
\begin{enumerate}[label=(\roman*)]
\item the \(Q_\Lambda\)-Ward identity has no boundary anomaly on
  \(\Gamma_\Lambda^\rho\);
\item \(Q_\Lambda\mathcal O_i=0\) for all insertions;
\item the integrated metric response has the form
\begin{equation}
  \delta_g S_\Lambda^\rho
  =
  Q_\Lambda \mathcal V_{\Lambda,\delta g}
  +
  \mathcal C_{\Lambda,\delta g},
  \label{eq:twist-metric-response}
\end{equation}
where \(\mathcal C_{\Lambda,\delta g}\) is the declared contact and anomaly
representative;
\item the insertion prescription makes
\[
  \left\langle
  \mathcal C_{\Lambda,\delta g}\prod_i\mathcal O_i
  \right\rangle_\Lambda=0 .
\]
\end{enumerate}
Then \(Z_g^\rho(\mathcal O_1,\ldots,\mathcal O_n)\) is locally constant in
the metric \(g\) at the regulator scale.  If the continuum limit exists and
the hypotheses are stable under that limit, the limiting correlator is
locally metric independent.
Differentiating the regulated path integral gives
\[
  \delta_g Z_g^\rho
  =
  -
  \left\langle
  \delta_g S_\Lambda^\rho\prod_i\mathcal O_i
  \right\rangle_\Lambda
  +
  \delta_g^{\rm contact}
  \left\langle\prod_i\mathcal O_i\right\rangle_\Lambda .
\]
Insert \eqref{eq:twist-metric-response}.  The term with
\(Q_\Lambda\mathcal V_{\Lambda,\delta g}\) is, up to the vanishing
\(Q_\Lambda\mathcal O_i\) terms, the \(Q_\Lambda\)-variation of
\(\mathcal V_{\Lambda,\delta g}\prod_i\mathcal O_i\).  The Ward identity
sets its expectation value to zero.  The remaining contact representative is
zero by hypothesis.  Hence the first variation vanishes.

The criterion is deliberately stated at regulator level.  A formal identity
\(\delta_gS=Q(\cdots)\) in a continuum Lagrangian is one ingredient; it
becomes a theorem about a QFT only after the regulated Ward identity,
contact terms, and boundary behavior are controlled.
\end{remark}

\section{\texorpdfstring{Four-Dimensional \(\mathcal N=2\) Supercharges}{Four-Dimensional N=2 Supercharges}}

In four Euclidean dimensions,
\[
  \operatorname{Spin}(4)=SU(2)_+\times SU(2)_- .
\]
The pure \(\mathcal N=2\) vector multiplet has \(R\)-symmetry
\[
  G_R=SU(2)_R\times U(1)_r
\]
at the classical level; the subgroup used in the Donaldson twist is
\(SU(2)_R\).  The supercharges transform as
\begin{equation}
  Q_\alpha^{\ A}\in (2,1;2)_{+1},
  \qquad
  \bar Q_{\dot\alpha A}\in (1,2;2)_{-1},
  \label{eq:n2-supercharge-representations}
\end{equation}
where \(\alpha\) is an \(SU(2)_+\) spinor index,
\(\dot\alpha\) is an \(SU(2)_-\) spinor index, \(A\) is an \(SU(2)_R\)
index, and the subscript denotes \(U(1)_r\) charge in the normalization used
in this chapter.  The Donaldson twisting homomorphism uses the \(SU(2)_R\)
factor.

The Donaldson twisting homomorphism identifies \(SU(2)_+\) with \(SU(2)_R\):
\[
  \rho_{\rm DW}:SU(2)_+\times SU(2)_-\longrightarrow SU(2)_R,
  \qquad
  \rho_{\rm DW}(s_+,s_-)=s_+ .
\]
The twisted rotation group is
\[
  \operatorname{Spin}(4)'=SU(2)'_+\times SU(2)_-,
  \qquad
  SU(2)'_+=\operatorname{diag}(SU(2)_+\times SU(2)_R).
\]

Under \(SU(2)'_+\times SU(2)_-\),
\[
  (2,1;2)
  \longmapsto
  (1,1)\oplus(3,1),
  \qquad
  (1,2;2)
  \longmapsto
  (2,2).
\]
Thus the twisted supercharges consist of a scalar \(Q\), a self-dual
two-form supercharge \(Q^+_{\mu\nu}\), and a one-form supercharge
\(Q_\mu\).
The first representation has an \(SU(2)_+\) doublet index and an \(SU(2)_R\)
doublet index.  Restriction to the diagonal \(SU(2)'_+\) gives
\[
  2\otimes 2=1\oplus3 .
\]
The singlet is represented by contraction with the invariant tensor
\(\epsilon^{\alpha A}\), hence
\[
  Q=\epsilon^{\alpha A}Q_{\alpha A}
\]
is a scalar.  The triplet is the symmetric traceless part and is identified
with self-dual two-forms because the adjoint representation of \(SU(2)_+\)
is \(\Lambda^{2,+}\).  The second representation has no \(SU(2)_+\) spinor
index; its \(SU(2)_R\) doublet becomes an \(SU(2)'_+\) doublet, while the
\(SU(2)_-\) doublet remains.  Therefore it becomes \((2,2)\), the vector
representation \(T^\ast M\).

\section{Donaldson-Witten Field Decomposition}

The untwisted vector multiplet contains a connection \(A_\mu\), complex
adjoint scalars \(\phi,\bar\phi\), gauginos
\(\lambda_\alpha^{\ A}\), \(\bar\lambda_{\dot\alpha A}\), and auxiliary
fields.  After the twist, the gauginos decompose by the same representation
calculation:
\[
  \lambda_\alpha^{\ A}
  \longmapsto
  \eta\oplus\chi^+_{\mu\nu},
  \qquad
  \bar\lambda_{\dot\alpha A}
  \longmapsto
  \psi_\mu .
\]
On a four-manifold \(M\) and a principal gauge bundle \(P\to M\), the
twisted fields are therefore
\[
  A\in\mathcal A(P),
  \qquad
  \phi,\bar\phi,\eta\in\Omega^0(M,\operatorname{ad}P),
\]
\[
  \psi\in\Omega^1(M,\operatorname{ad}P),
  \qquad
  \chi^+,H^+\in\Omega^{2,+}(M,\operatorname{ad}P).
\]
The fields \(\eta,\psi,\chi^+\) are odd; \(A,\phi,\bar\phi,H^+\) are even.
The auxiliary field \(H^+\) is included so that the scalar symmetry closes
off shell.

Use the infinitesimal gauge convention
\begin{equation}
  \delta_\epsilon A=d_A\epsilon,
  \qquad
  \delta_\epsilon X=-[\epsilon,X]
  \quad\text{for }X\in\Omega^\bullet(M,\operatorname{ad}P).
  \label{eq:twist-gauge-convention}
\end{equation}
The Donaldson-Witten scalar supercharge acts by
\begin{equation}
\begin{aligned}
  &QA=\psi,\qquad &&Q\psi=-d_A\phi,\qquad &&Q\phi=0,\\
  &Q\bar\phi=\eta,\qquad &&Q\eta=[\phi,\bar\phi],\\
  &Q\chi^+=H^+,\qquad &&QH^+=[\phi,\chi^+] .
\end{aligned}
\label{eq:donaldson-witten-q-transformations}
\end{equation}

\paragraph{Off-shell closure.}
With the convention \eqref{eq:twist-gauge-convention}, the transformations
\eqref{eq:donaldson-witten-q-transformations} satisfy
\[
  Q^2=\delta_{-\phi}
\]
on all twisted vector-multiplet fields.  Hence \(Q^2=0\) on gauge-invariant
observables.

For the connection,
\[
  Q^2A=Q\psi=-d_A\phi=\delta_{-\phi}A.
\]
For an adjoint field \(X\), \(\delta_{-\phi}X=[\phi,X]\).  The transformations
give
\[
  Q^2\bar\phi=Q\eta=[\phi,\bar\phi],
  \qquad
  Q^2\chi^+=QH^+=[\phi,\chi^+],
  \qquad
  Q^2\phi=0=[\phi,\phi].
\]
It remains to check \(\psi\) and \(H^+\).  Since \(QA=\psi\) and
\(Q\phi=0\),
\[
  Q(d_A\phi)=[\psi,\phi],
\]
where the bracket includes the wedge product of forms and the Lie bracket in
\(\operatorname{ad}P\).  Hence
\[
  Q^2\psi=-[\psi,\phi]=[\phi,\psi]=\delta_{-\phi}\psi.
\]
Finally,
\[
  Q^2H^+=Q[\phi,\chi^+]=[\phi,H^+]=\delta_{-\phi}H^+ .
\]
Thus the closure relation holds on each generator and therefore on the
graded algebra they generate.

\section{The Donaldson-Witten Gauge Fermion}

Let the monograph gauge trace convention be
\(\operatorname{tr}(t^at^b)=\delta^{ab}\).  The Yang-Mills action is written
with
\[
  \frac{1}{4g_{\rm YM}^2}
  \int_M d{\rm vol}_g\,\operatorname{tr}(F_{\mu\nu}F^{\mu\nu})
  =
  \frac{1}{2g_{\rm YM}^2}
  \int_M\operatorname{tr}(F\wedge *F).
\]
The topological term uses
\[
  k(P)=-\frac{1}{8\pi^2}\int_M\operatorname{tr}(F_A\wedge F_A).
\]

A convenient cohomological representative of the Donaldson-Witten action has
the form
\begin{equation}
  S_{\rm DW}
  =
  \frac{1}{g_{\rm YM}^2}Q\Psi
  -\frac{i\theta}{8\pi^2}\int_M\operatorname{tr}(F_A\wedge F_A),
  \label{eq:donaldson-witten-action-q-exact}
\end{equation}
where the gauge fermion contains the Mathai-Quillen term
\begin{equation}
  \Psi_{\rm MQ}
  =
  \int_M
  \operatorname{tr}\left[
  \chi^+\wedge *
  \left(F_A^+ +\frac12 H^+\right)
  \right]
  +\text{terms fixing the scalar/gauge directions}.
  \label{eq:dw-mq-gauge-fermion}
\end{equation}
Applying \(Q\) to the displayed term gives
\[
  Q\Psi_{\rm MQ}
  =
  \int_M\operatorname{tr}\left[
  H^+\wedge *\left(F_A^++\frac12H^+\right)
  -
  \chi^+\wedge *d_A^+\psi
  \right],
\]
up to the sign convention for moving the odd field \(\chi^+\) through \(Q\).
The algebraic equation of motion for \(H^+\) is
\[
  H^+=-F_A^+ .
\]
Substitution gives the bosonic square
\[
  \frac12\int_M\operatorname{tr}(F_A^+\wedge *F_A^+)
\]
with the convention of \eqref{eq:dw-mq-gauge-fermion}.  Rescaling
\(\Psi_{\rm MQ}\) rescales this square; the localization locus is the
zero-set of the section
\[
  F_A^+=0.
\]

The cohomological fields \((\psi,\chi^+,H^+)\) are precisely the
gauge-equivariant Mathai-Quillen variables for the section
\[
  F^+:\mathcal A(P)\longrightarrow
  \Omega^{2,+}(M,\operatorname{ad}P).
\]
After quotienting by the gauge group, the virtual tangent-obstruction
complex at an anti-self-dual connection is
\[
  \Omega^0(M,\operatorname{ad}P)
  \xrightarrow{d_A}
  \Omega^1(M,\operatorname{ad}P)
  \xrightarrow{d_A^+}
  \Omega^{2,+}(M,\operatorname{ad}P),
\]
which is the complex used in the Witten-Donaldson chapter to compute the
expected dimension.

\section{Donaldson Observables from Descent}

Let \(p\) be an invariant polynomial on \(\mathfrak g\).  For
\(\mathfrak g=\mathfrak{su}(2)\) the basic example in the monograph trace
convention is
\[
  p(\phi)=\frac12\operatorname{tr}(\phi^2).
\]
Using the universal curvature combination
\[
  \mathcal F_Q
  =
  F_A+\psi+\phi
\]
graded by form degree plus \(Q\)-degree, the Bianchi identity and
\eqref{eq:donaldson-witten-q-transformations} imply the equivariant closure
\[
  (Q+d_A)\mathcal F_Q=0
  \quad\text{modulo the gauge action}.
\]
Consequently
\[
  (Q+d)\operatorname{tr}(\mathcal F_Q^2)=0
\]
on gauge-invariant expressions.  Writing
\[
  \frac12\operatorname{tr}(\mathcal F_Q^2)
  =
  \mathcal O^{(0)}+\mathcal O^{(1)}+\mathcal O^{(2)}
  +\mathcal O^{(3)}+\mathcal O^{(4)}
\]
by form degree gives the descent equations
\[
  Q\mathcal O^{(0)}=0,
  \qquad
  Q\mathcal O^{(r)}+d\mathcal O^{(r-1)}=0
  \quad(1\le r\le4).
\]
For a closed \(r\)-cycle \(\gamma\subset M\),
\[
  \mathcal O_\gamma=\int_\gamma\mathcal O^{(r)}
\]
is \(Q\)-closed.  If \(\gamma-\gamma'=\partial C\), then
\[
  \mathcal O_\gamma-\mathcal O_{\gamma'}
  =
  -Q\int_C\mathcal O^{(r+1)}.
\]
Thus Donaldson observables depend on homology classes in \(Q\)-cohomology,
subject to the contact-term prescription for products of integrated
descendants.

\section{\texorpdfstring{Two-Dimensional \(\mathcal N=(2,2)\) A and B Twists}{Two-Dimensional N=(2,2) A and B Twists}}

Let the Euclidean rotation group in two dimensions be \(U(1)_{\rm rot}\).
Use supercharges
\[
  Q_+,\quad Q_-,\quad \bar Q_+,\quad \bar Q_- ,
\]
with spins
\[
  s(Q_+)=s(\bar Q_+)=+\frac12,
  \qquad
  s(Q_-)=s(\bar Q_-)=-\frac12 .
\]
Assume both vector and axial \(R\)-symmetries exist as nonanomalous
symmetries.  Fix charges
\[
\begin{array}{c|rrrr}
 & Q_+ & Q_- & \bar Q_+ & \bar Q_-\\ \hline
q_V & +1 & +1 & -1 & -1\\
q_A & +1 & -1 & -1 & +1 .
\end{array}
\]
With the convention
\[
  s_\rho=s+\frac12 q_R ,
\]
the A-twist uses \(q_R=q_V\) and has scalar supercharges
\[
  \bar Q_+,\qquad Q_- .
\]
The B-twist uses \(q_R=q_A\) and has scalar supercharges
\[
  \bar Q_+,\qquad \bar Q_- .
\]
The standard BRST choices are
\[
  Q_A=\bar Q_+ + Q_-,
  \qquad
  Q_B=\bar Q_+ + \bar Q_- .
\]
The A-model and B-model complexes developed in
Chapter~\ref{ch:tqft-topological-sigma-models} arise from these two scalar
supercharges after twisting the \(\mathcal N=(2,2)\) sigma-model fields.
The vector or axial \(R\)-symmetry used in the twist must be nonanomalous for
the chosen target and background.  In sigma models this condition becomes
the familiar restriction: the A-twist is controlled by symplectic data, while
the closed B-twist requires the Calabi-Yau condition for the axial
\(R\)-symmetry and trace pairing.

\section{What the Twist Does and What Remains to Construct}

The twist provides a representation-theoretic and geometric conversion:
\[
  \text{spinors with \(R\)-indices}
  \longrightarrow
  \text{forms or tensors on \(M\)}.
\]
It also identifies candidate scalar supercharges and candidate descent
complexes.  A theorem-level cohomological QFT requires the additional
regulator-level closure statement \eqref{eq:twist-q-square-general} and
Remark~\ref{rem:twist-metric-independence}: \(Q\)-invariant
regularization, anomaly cancellation, off-shell closure or BV closure,
contact representatives, compactification of fixed loci, and gluing.

\begin{openproblem}[Twisted continuum construction]
\label{op:twisted-continuum-construction}
Construct four-dimensional twisted supersymmetric gauge theories as
continuum QFTs with \(Q\)-preserving regulator or BV regularization,
off-shell or BV closure, determinant-line and global-form control, contact
term prescriptions for descendant observables, compactified fixed loci, and
a derivation of the \(Q\)-exact metric-response relation.  For
Donaldson-Witten theory this construction should recover the
anti-self-dual-moduli-space description and the Donaldson observables from
the regulated cohomological gauge theory.
\end{openproblem}
