\chapter{Bordism Functoriality and Extended Topological Field Theory}
\label{ch:tqft-bordism-functoriality-extended}

\ConventionsEuclidean

The cohomological mechanism of
Chapter~\ref{ch:tqft-metric-independence} produces correlation functionals
whose metric variation is controlled by a Ward identity.  A functorial
topological field theory is a stronger object.  It assigns state spaces to
closed spatial manifolds, linear maps to bordisms, and compatibility maps for
cutting and gluing.  The purpose of this chapter is to state the functorial
data precisely enough that later claims about BF theory, Chern--Simons
theory, finite gauge theory, cohomological gauge theory, and topological
twists have a definite target.

Throughout this chapter \(\Bbbk\) is a field of characteristic zero unless a
finite-field example is explicitly declared.  The finite-dimensional target
is denoted \({\bf Vect}^{\rm fd}_{\Bbbk}\).  Infinite-dimensional state
spaces, bornological vector spaces, and Hilbert completions occur naturally in
non-topological QFT and in some noncompact topological theories; whenever
they are used later, the topology on the target is part of the datum and the
gluing maps must be continuous for that topology.

\section{Tangential Structures and Collared Bordisms}

A tangential structure in dimension \(D\) is a map
\[
  \theta_\xi:B_\xi\longrightarrow BO(D).
\]
Examples are \(BSO(D)\to BO(D)\) for oriented manifolds,
\(BSpin(D)\to BO(D)\) for spin manifolds, and the point
\(\ast\to BO(D)\) classifying a framing of the tangent bundle.  If \(N\) is a
smooth \(d\)-manifold with \(d\le D\), a \(\xi\)-structure on \(N\) means a
lift, up to homotopy through lifts, of the classifying map of the stabilized
tangent bundle
\[
  TN\oplus\mathbb R^{D-d}\longrightarrow BO(D)
\]
to \(B_\xi\).  This stabilized convention is the one compatible with
extended bordisms: a hypersurface, a boundary corner, and a point are all
regarded as lower-dimensional strata inside \(D\)-dimensional bordisms.

\begin{definition}[Collared \(\xi\)-bordism]
\label{def:collared-xi-bordism}
Let \(\Sigma_0\) and \(\Sigma_1\) be closed compact smooth
\((D-1)\)-manifolds with \(\xi\)-structure.  A collared \(\xi\)-bordism
\[
  M:\Sigma_0\longrightarrow\Sigma_1
\]
is a compact smooth \(D\)-manifold \(M\) with \(\xi\)-structure, together
with boundary identifications
\[
  \partial M\simeq \bar\Sigma_0\sqcup\Sigma_1
\]
and collar embeddings near the incoming and outgoing components.  Here
\(\bar\Sigma_0\) denotes the same underlying manifold with the tangential
structure reversed in the sense appropriate to \(\xi\).  Two collared
bordisms are equivalent if they are related by a \(\xi\)-diffeomorphism
which respects the collars up to isotopy.
\end{definition}

The collars are not decorative.  They make composition a smooth operation:
if \(M_{01}:\Sigma_0\to\Sigma_1\) and \(M_{12}:\Sigma_1\to\Sigma_2\), then
\[
  M_{12}\circ M_{01}
  =
  M_{01}\cup_{\Sigma_1}M_{12}
\]
is obtained by identifying the outgoing collar of \(M_{01}\) with the
incoming collar of \(M_{12}\), followed by the canonical smoothing determined
by the collar product coordinate.  The identity of \(\Sigma\) is the
cylinder \(\Sigma\times[0,1]\) with product \(\xi\)-structure.

\begin{definition}[Ordinary bordism category]
\label{def:ordinary-bordism-category}
The category \(\operatorname{Bord}^{\xi}_{D}\) has:
\begin{itemize}
\item objects: closed compact smooth \((D-1)\)-manifolds with
  \(\xi\)-structure;
\item morphisms: equivalence classes of collared \(\xi\)-bordisms;
\item composition: collar gluing;
\item symmetric monoidal product: disjoint union.
\end{itemize}
The monoidal unit is the empty manifold \(\varnothing\).
\end{definition}

Figure~\ref{fig:tqft-collar-gluing} records the collar-gluing operation in the
bordism category: two bordisms sharing the collared boundary \(\Sigma_1\)
compose by identifying that boundary.

\begin{figure}[H]
\centering
\begin{tikzpicture}[>=Stealth,scale=0.95]
  \draw[thick] (0,0) rectangle (2.4,1.2);
  \draw[thick] (2.4,0) rectangle (4.8,1.2);
  \draw[very thick,blue!55!black] (2.4,0) -- (2.4,1.2);
  \node at (1.2,0.6) {\(M_{01}\)};
  \node at (3.6,0.6) {\(M_{12}\)};
  \node[below] at (0,0) {\(\Sigma_0\)};
  \node[below] at (2.4,0) {\(\Sigma_1\)};
  \node[below] at (4.8,0) {\(\Sigma_2\)};
  \draw[->] (5.25,0.6) -- (6.15,0.6);
  \draw[thick] (6.6,0) rectangle (9.4,1.2);
  \node at (8.0,0.6) {\(M_{12}\circ M_{01}\)};
  \node[below] at (6.6,0) {\(\Sigma_0\)};
  \node[below] at (9.4,0) {\(\Sigma_2\)};
\end{tikzpicture}
\caption{Composition in the ordinary bordism category is collar gluing.  The
intermediate boundary component is not retained as extra data after the
composition; extended TQFT keeps track of such cuts through lower-codimension
assignments.}
\label{fig:tqft-collar-gluing}
\end{figure}

Definition~\ref{def:ordinary-bordism-category} is the \(1\)-categorical
version.  A fully smooth functorial QFT normally uses a topologically
enriched or higher bordism object in which smooth families of bordisms are
also visible.  The ordinary category is nevertheless the right first target
for finite-dimensional Atiyah--Segal TQFT because it isolates the exact
gluing identities on which every refinement must reduce.

\section{Atiyah--Segal Functoriality}

\begin{definition}[Atiyah--Segal topological field theory]
\label{def:atiyah-segal-tqft}
An honest \(D\)-dimensional Atiyah--Segal TQFT with tangential structure
\(\xi\) and finite-dimensional linear target is a symmetric monoidal functor
\[
  Z:\operatorname{Bord}^{\xi}_{D}
  \longrightarrow
  {\bf Vect}^{\rm fd}_{\Bbbk}.
\]
Thus there are canonical isomorphisms
\[
  Z(\Sigma_1\sqcup\Sigma_2)
  \simeq
  Z(\Sigma_1)\otimes Z(\Sigma_2),
  \qquad
  Z(\varnothing)\simeq \Bbbk,
\]
and for composable bordisms
\[
  Z(M_{12}\circ M_{01})=Z(M_{12})\,Z(M_{01}).
\]
The cylinder maps to the identity on \(Z(\Sigma)\).
\end{definition}

For a closed \(D\)-manifold \(M\), regarded as a bordism
\(\varnothing\to\varnothing\), the value \(Z(M)\) is the scalar by which the
map \(\Bbbk\to\Bbbk\) multiplies.  For a bordism
\(M:\Sigma_0\to\Sigma_1\), \(Z(M)\) is the transition amplitude from the
state space \(Z(\Sigma_0)\) to \(Z(\Sigma_1)\).  These words are justified by
the gluing axiom, not by an analogy with time evolution: if \(M\) is cut
along an intermediate closed hypersurface \(\Sigma\), the amplitude factors
through the state space assigned to \(\Sigma\).

\begin{definition}[Anomalous or relative functorial theory]
\label{def:anomalous-relative-tqft}
Let \(\mathcal L\) be an invertible \((D+1)\)-dimensional field theory, or
equivalently in the finite-dimensional setting a compatible system of
one-dimensional anomaly lines over boundary data.  A \(D\)-dimensional theory
relative to \(\mathcal L\) assigns vectors or linear maps valued in those
lines.  It becomes an honest functor only after the anomaly line has been
trivialized, or after the \((D+1)\)-dimensional bulk \(\mathcal L\) is kept
as part of the system.
\end{definition}

This distinction matters in gauge theory and in theories with gravitational
or global anomalies.  A path integral may satisfy topological Ward identities
on closed manifolds while its boundary Hilbert spaces transform projectively.
The functorial object is then relative, not an honest functor
\(\operatorname{Bord}^{\xi}_{D}\to{\bf Vect}_{\Bbbk}^{\rm fd}\).

The Atiyah--Segal datum is the metric-independent specialization of the
Kontsevich--Segal viewpoint developed in
Chapter~\ref{ch:kontsevich-segal-functorial-qft}.  A
Kontsevich--Segal theory keeps admissible complex metrics and analytic
dependence as part of the bordism datum.  An Atiyah--Segal TQFT is obtained
when, after all anomaly and background-structure data have been included, the
amplitude is locally constant on the geometric parameters.

\section{Duality, Pairings, and Reflection}

If \(\bar\Sigma\) is the reversed \(\xi\)-manifold, the evaluation bordism
\[
  {\rm ev}_\Sigma:\bar\Sigma\sqcup\Sigma\longrightarrow\varnothing
\]
and coevaluation bordism
\[
  {\rm coev}_\Sigma:\varnothing\longrightarrow
  \Sigma\sqcup\bar\Sigma
\]
are the two caps obtained by viewing the cylinder on \(\Sigma\) sideways.
Functoriality gives maps
\[
  Z({\rm ev}_\Sigma):
  Z(\bar\Sigma)\otimes Z(\Sigma)\to\Bbbk,
  \qquad
  Z({\rm coev}_\Sigma):
  \Bbbk\to Z(\Sigma)\otimes Z(\bar\Sigma).
\]

\begin{proposition}[State-space duality from caps and cylinders]
\label{prop:tqft-state-space-duality}
Let \(Z\) be an honest finite-dimensional Atiyah--Segal TQFT.  Then
\(Z(\bar\Sigma)\) is canonically dual to \(Z(\Sigma)\).  Equivalently, the
pairing \(Z(\bar\Sigma)\otimes Z(\Sigma)\to\Bbbk\) from the evaluation
bordism is nondegenerate, and the coevaluation vector is its inverse tensor.
\end{proposition}

\begin{proof}
The two zig-zag composites
\[
  \Sigma
  \xrightarrow{\ {\rm coev}_\Sigma\sqcup\id_\Sigma\ }
  \Sigma\sqcup\bar\Sigma\sqcup\Sigma
  \xrightarrow{\ \id_\Sigma\sqcup{\rm ev}_\Sigma\ }
  \Sigma
\]
and
\[
  \bar\Sigma
  \xrightarrow{\ \id_{\bar\Sigma}\sqcup{\rm coev}_\Sigma\ }
  \bar\Sigma\sqcup\Sigma\sqcup\bar\Sigma
  \xrightarrow{\ {\rm ev}_\Sigma\sqcup\id_{\bar\Sigma}\ }
  \bar\Sigma
\]
are diffeomorphic to the corresponding cylinders.  Applying \(Z\), and using
the cylinder axiom, gives the triangular identities for a dual pair in
\({\bf Vect}^{\rm fd}_{\Bbbk}\).  In finite dimension those triangular
identities are equivalent to nondegeneracy of the evaluation pairing and to
the assertion that the coevaluation vector is the inverse of that pairing.
\end{proof}

A unitary topological theory requires more than the bilinear pairing above.
Reflection of bordisms should be implemented by an antilinear involution on
state spaces, and the pairing obtained by reflecting a state and gluing it to
the original state should be positive.  This is the TQFT analogue of
reflection positivity in Euclidean QFT.  Positivity is therefore not a
consequence of Definition~\ref{def:atiyah-segal-tqft}; it is extra structure
on the functor.

\section{The Complete Two-Dimensional Oriented Example}

The first nontrivial test of the definition is the classification of ordinary
two-dimensional oriented TQFTs by finite-dimensional commutative Frobenius
algebras.  This theorem gives the exact algebraic form of the gluing law for
surfaces, with pair-of-pants multiplication as one piece of the structure.

\begin{definition}[Commutative Frobenius algebra]
\label{def:commutative-frobenius-algebra}
A commutative Frobenius algebra over \(\Bbbk\) is a finite-dimensional
unital associative commutative \(\Bbbk\)-algebra \(A\), together with a
linear functional
\[
  \epsilon:A\to\Bbbk
\]
such that the bilinear form
\begin{equation}
  \eta(a,b)=\epsilon(ab)
  \label{eq:frobenius-pairing}
\end{equation}
is nondegenerate.  The unit is denoted \(1_A\), the multiplication is
\(m:A\otimes A\to A\), and the counit is \(\epsilon\).
\end{definition}

Choose a basis \(\{e_i\}_{i=1}^{N}\) of \(A\).  Let
\[
  \eta_{ij}=\epsilon(e_i e_j),
  \qquad
  (\eta^{ij})=(\eta_{ij})^{-1},
\]
and define the dual elements
\[
  e^i=\sum_j \eta^{ij}e_j .
\]
Then \(\epsilon(e_i e^j)=\delta_i^{\,j}\).  The inverse of the Frobenius
pairing is the tensor
\[
  C=\sum_i e_i\otimes e^i\in A\otimes A.
\]
The comultiplication adjoint to multiplication is
\begin{equation}
  \Delta(a)
  =
  \sum_i e_i\otimes e^i a
  =
  \sum_i ae_i\otimes e^i .
  \label{eq:frobenius-comultiplication}
\end{equation}
The equality of the two expressions follows from commutativity.  The defining
adjointness property is
\begin{equation}
  (\epsilon\otimes\epsilon)\bigl(\Delta(a)(b\otimes c)\bigr)
  =
  \epsilon(abc),
  \qquad a,b,c\in A.
  \label{eq:delta-adjoint-to-m}
\end{equation}

\begin{lemma}[Cylinder and Frobenius identities]
\label{lem:frobenius-cylinder-and-neck}
For a commutative Frobenius algebra \(A\), the tensor \(C=\sum_i e_i\otimes
e^i\) and the comultiplication \(\Delta\) satisfy:
\begin{align}
  \sum_i e_i\,\epsilon(e^i a)&=a,
  \label{eq:frobenius-cylinder-identity}\\
  (m\otimes\id_A)(\id_A\otimes\Delta)
  &=
  \Delta m
  =
  (\id_A\otimes m)(\Delta\otimes\id_A).
  \label{eq:frobenius-neck-identity}
\end{align}
\end{lemma}

\begin{proof}
For any \(b\in A\),
\[
  \epsilon\!\left(b\sum_i e_i\,\epsilon(e^i a)\right)
  =
  \sum_i \epsilon(be_i)\epsilon(e^i a)
  =
  \epsilon(ba),
\]
because \(\{e_i\}\) and \(\{e^i\}\) are dual bases for the pairing
\(\eta\).  Nondegeneracy of \(\eta\) gives
\eqref{eq:frobenius-cylinder-identity}.

For the first equality in \eqref{eq:frobenius-neck-identity}, evaluate both
sides on \(a\otimes b\).  The left side is
\[
  (m\otimes\id)(a\otimes\Delta(b))
  =
  \sum_i ae_i\otimes e^i b .
\]
The middle term is
\[
  \Delta(ab)=\sum_i e_i\otimes e^i ab .
\]
To prove equality, pair the first tensor factor with an arbitrary \(x\in A\)
and the second with an arbitrary \(y\in A\).  The two pairings give
\[
  \sum_i \epsilon(xae_i)\epsilon(ye^i b)
  =
  \epsilon(xyab)
  =
  \sum_i \epsilon(xe_i)\epsilon(ye^i ab),
\]
again by the dual-basis identity.  Nondegeneracy of
\(\eta\otimes\eta\) proves the equality.  The second equality is identical,
or follows by commutativity.
\end{proof}

\begin{theorem}[Oriented \(2\)-dimensional TQFTs and Frobenius algebras]
\label{thm:2d-oriented-tqft-frobenius}
Finite-dimensional ordinary oriented \(2\)-dimensional Atiyah--Segal TQFTs
over \(\Bbbk\) are equivalent to finite-dimensional commutative Frobenius
algebras over \(\Bbbk\).  Under this equivalence,
\[
  A=Z(S^1),
  \qquad
  m=Z(P_{2\to1}),
  \qquad
  1_A=Z(D_{\varnothing\to1})(1),
  \qquad
  \epsilon=Z(D_{1\to\varnothing}),
\]
where \(P_{2\to1}\) is the oriented pair of pants with two incoming circles
and one outgoing circle.
\end{theorem}

\begin{proof}
First let \(Z\) be a \(2\)-dimensional oriented TQFT and set
\(A=Z(S^1)\).  The disk from \(\varnothing\) to \(S^1\) gives the unit
\(1_A\), the reversed disk gives \(\epsilon:A\to\Bbbk\), and the pair of
pants gives \(m:A\otimes A\to A\).  Associativity follows because the two
bordisms
\[
  P_{2\to1}\circ(P_{2\to1}\sqcup\id_{S^1})
  \quad\hbox{and}\quad
  P_{2\to1}\circ(\id_{S^1}\sqcup P_{2\to1})
\]
are diffeomorphic as oriented bordisms from three circles to one circle.
Commutativity follows from the diffeomorphism of the pair of pants which
interchanges the two incoming circles.  The unit axiom follows by gluing a
disk to one incoming circle of the pair of pants, producing a cylinder.
Proposition~\ref{prop:tqft-state-space-duality}, applied to the circle,
shows that \(\epsilon(ab)\) is nondegenerate.  Thus \(A\) is a commutative
Frobenius algebra.

Conversely, let \((A,\epsilon)\) be a commutative Frobenius algebra.  Define
the value on a disjoint union of \(n\) positively oriented circles to be
\(A^{\otimes n}\); reversed circles are assigned the dual vector space, and
the Frobenius pairing identifies it with \(A\) when an explicit formula is
written using the basis-independent tensor \(C\).  Assign to the elementary
Morse bordisms the maps
\[
  D_{\varnothing\to1}\mapsto 1_A,\qquad
  D_{1\to\varnothing}\mapsto\epsilon,\qquad
  P_{2\to1}\mapsto m,\qquad
  P_{1\to2}\mapsto\Delta .
\]
For a general compact oriented surface with collared boundary, choose a
Morse function constant near the boundary.  The critical points decompose
the surface into disks and pairs of pants, hence into a composite of the
elementary maps above.

It remains to show that the resulting linear map is independent of the Morse
decomposition.  The differential-topological input is the standard Cerf
description of decompositions of compact surfaces: two decompositions are
related by isotopy, by creation or cancellation of adjacent critical points,
and by interchange of neighboring critical levels.  Isotopy gives no change.
Cancellation of a birth-death pair gives precisely the cylinder identity
\eqref{eq:frobenius-cylinder-identity} and the unit/counit identities.
Interchange of two pair-of-pants critical levels gives associativity,
commutativity, and the neck-exchange Frobenius identity
\eqref{eq:frobenius-neck-identity}.  Therefore the map depends only on the
oriented bordism class, and collar gluing is sent to composition because
concatenating Morse functions concatenates the associated elementary maps.
The construction is symmetric monoidal by tensor product on disjoint unions.

The two constructions are inverse up to the evident natural isomorphism: a
TQFT is generated by disks and pairs of pants modulo the relations just
checked, and a Frobenius algebra is recovered from the values of the
constructed functor on those generating bordisms.
\end{proof}

Figure~\ref{fig:tqft-frobenius-generators} displays the generating bordisms for
the two-dimensional oriented TQFT/Frobenius-algebra equivalence: multiplication,
unit, pairing, and the relations imposed by bordism invariance.

\begin{figure}[H]
\centering
\begin{tikzpicture}[>=Stealth,scale=0.92]
  \draw[thick] (0,0.8) .. controls (0.9,0.8) and (1.2,0.35) .. (1.9,0.35)
               .. controls (2.6,0.35) and (3.0,0.8) .. (3.9,0.8);
  \draw[thick] (0,-0.8) .. controls (0.9,-0.8) and (1.2,-0.35) .. (1.9,-0.35)
               .. controls (2.6,-0.35) and (3.0,-0.8) .. (3.9,-0.8);
  \draw[thick] (1.9,0.35) .. controls (2.1,0.1) and (2.1,-0.1) .. (1.9,-0.35);
  \node[left] at (0,0.8) {\(a\)};
  \node[left] at (0,-0.8) {\(b\)};
  \node[right] at (3.9,0) {\(m(a,b)\)};
  \node at (1.95,-1.25) {pair of pants \(P_{2\to1}\)};

  \draw[->] (4.8,0) -- (5.7,0);

  \draw[thick] (6.2,0) ellipse (0.45 and 0.85);
  \draw[thick] (7.1,0.55) .. controls (7.8,0.55) and (8.1,0.2) .. (8.7,0.2);
  \draw[thick] (7.1,-0.55) .. controls (7.8,-0.55) and (8.1,-0.2) .. (8.7,-0.2);
  \draw[thick] (8.7,0.2) .. controls (8.95,0.05) and (8.95,-0.05) .. (8.7,-0.2);
  \node at (6.2,-1.25) {cap/cup tensor \(C=\sum_i e_i\otimes e^i\)};
  \node at (8.25,-1.25) {comultiplication \(\Delta\)};
\end{tikzpicture}
\caption{In two dimensions the gluing generators are disks, caps, cups, and
pairs of pants.  The inverse pairing tensor \(C\) is what makes gluing along
an internal circle a contraction over the intermediate state space.}
\label{fig:tqft-frobenius-generators}
\end{figure}

\begin{example}[Semisimple idempotent basis]
\label{ex:semisimple-2d-tqft}
Let
\[
  A=\Bbbk^r
\]
with pointwise multiplication.  Let \(p_\alpha\), \(1\le\alpha\le r\), be the
primitive idempotents and choose nonzero weights
\(\lambda_\alpha\in\Bbbk^\times\).  Define
\[
  \epsilon(p_\alpha)=\lambda_\alpha .
\]
Then
\[
  \eta(p_\alpha,p_\beta)=\lambda_\alpha\delta_{\alpha\beta},
  \qquad
  p^\alpha=\lambda_\alpha^{-1}p_\alpha,
\]
and the Euler element
\[
  E=m\Delta(1_A)=\sum_{\alpha=1}^{r}\lambda_\alpha^{-1}p_\alpha
\]
is the handle operator.  For a closed connected oriented surface
\(\Sigma_g\) of genus \(g\),
\begin{equation}
  Z(\Sigma_g)=\epsilon(E^g)
  =
  \sum_{\alpha=1}^{r}\lambda_\alpha^{1-g}.
  \label{eq:semisimple-2d-tqft-genus-partition}
\end{equation}
The formula is obtained by cutting \(\Sigma_g\) into a sphere with \(g\)
handles.  Each handle inserts \(E\); the final cap applies \(\epsilon\).
\end{example}

\section{Extended Theories and What Extra Data They Contain}

An extended TQFT assigns data not only to closed \((D-1)\)-manifolds but also
to lower-dimensional strata.  Its source is an extended bordism
\((\infty,D)\)-category
\[
  \operatorname{Bord}^{\xi,{\rm ext}}_{D},
\]
whose objects are points with \(\xi\)-structure, whose \(1\)-morphisms are
collared intervals and \(1\)-manifolds, whose \(2\)-morphisms are surfaces
with corners, and so on up to \(D\)-morphisms.  The target is a symmetric
monoidal higher category \(\mathcal C\).  A fully extended TQFT is a
symmetric monoidal higher functor
\[
  Z:\operatorname{Bord}^{\xi,{\rm ext}}_{D}\to\mathcal C .
\]
The phrase ``locality at all codimensions'' means that cutting along corners,
defects, junctions, and points is represented by composition in
\(\mathcal C\).

\begin{quotedtheorem}[Cobordism hypothesis, framed fully extended form]
\label{thm:cobordism-hypothesis-boundary}
Let \(\mathcal C\) be a symmetric monoidal \((\infty,D)\)-category with the
duals and adjoints needed to evaluate framed bordisms.  An object
\(X\in\mathcal C\) is fully \(D\)-dualizable when \(X\) has a dual, the
evaluation and coevaluation \(1\)-morphisms admit left and right adjoints,
those adjunction data admit the required adjoints one categorical level
higher, and this iterated adjointability continues through level \(D\).
Under these hypotheses, fully extended framed \(D\)-dimensional topological
field theories with target \(\mathcal C\) are classified by fully
\(D\)-dualizable objects of \(\mathcal C\), with equivalence understood as
equivalence of symmetric monoidal higher functors.
\end{quotedtheorem}

Theorem~\ref{thm:cobordism-hypothesis-boundary} is quoted as a boundary
statement.  Its proof is not reproduced in this monograph.  Its role here is
to fix the kind of target data that must be supplied when an example is said
to be fully extended: the target higher category, the object assigned to a
point, the required duals and adjoints, the tangential-structure action, and
any anomaly or boundary decoration.

\begin{example}[The extended refinement of the semisimple \(2\)-dimensional model]
\label{ex:extended-semisimple-2d-tqft}
Let \({\bf Alg}_2^{\rm fd}\) be the Morita \(2\)-category whose objects are
finite-dimensional \(\Bbbk\)-algebras, whose \(1\)-morphisms are finite
bimodules, and whose \(2\)-morphisms are bimodule maps.  A finite-dimensional
separable algebra \(A\) is fully dualizable in this \(2\)-category.  A
separability idempotent is an element
\[
  e=\sum_r x_r\otimes y_r\in A\otimes A^{\rm op}
\]
such that
\[
  \sum_r x_ry_r=1_A,
  \qquad
  \sum_r ax_r\otimes y_r=\sum_r x_r\otimes y_ra
  \quad\hbox{for all }a\in A.
\]
These two identities give the coevaluation and evaluation bimodules their
adjoints.  If \(A\) also has a symmetric nondegenerate trace
\(\epsilon:A\to\Bbbk\), then the oriented structure is supplied by the
Calabi--Yau trace.  The value on the circle is the Hochschild trace
\[
  HH_0(A)=A/[A,A].
\]
For the commutative semisimple algebra \(A=\Bbbk^r\), this is \(A\) itself,
and the ordinary theory obtained by forgetting point and interval data is the
Frobenius-algebra TQFT of
Example~\ref{ex:semisimple-2d-tqft}.
\end{example}

This example illustrates the conceptual difference between an ordinary and a
fully extended theory.  In the ordinary theory the circle state space is the
primitive datum.  In the fully extended Morita-theoretic description the
point value is the algebra \(A\), the interval value is the regular
bimodule, and the circle state space is obtained by taking the trace of the
identity \(A\)-bimodule.

\section{From Local QFT to Functorial Data}

A local QFT with a protected topological sector can produce a functorial
TQFT only after several independent statements have been proved.  The
cohomological Ward identity must show metric independence of the selected
observables.  The state spaces obtained by quantizing on closed spatial
manifolds must be identified and completed.  Cutting a spacetime along a
hypersurface must give an actual contraction over the intermediate state
space.  Gauge redundancies must be treated by a BV or lattice construction
which supplies a gauge-independent integral and a boundary-condition
category.  Finally, possible anomaly lines must either be trivialized or kept
as relative data.

\paragraph{Functorial extraction criterion.}
Let a regulated local QFT assign, for each collared \(\xi\)-bordism
\(M:\Sigma_0\to\Sigma_1\), a multilinear amplitude
\[
  Z_{\rm reg}(M):\mathcal H_{\rm reg}(\Sigma_0)\to
  \mathcal H_{\rm reg}(\Sigma_1)
\]
depending on a regulator and auxiliary geometric choices.  Suppose:
\begin{enumerate}[label=(\roman*)]
\item the selected state spaces have regulator-independent finite-dimensional
  limits \(\mathcal H(\Sigma)\);
\item for every smooth family of metrics or auxiliary choices, the
  renormalized variation of the selected amplitudes is zero or is valued in a
  specified anomaly line;
\item for every decomposition \(M=M_{12}\circ M_{01}\), the regulated gluing
  identity converges to
  \[
    Z(M)=Z(M_{12})Z(M_{01})
  \]
  in the operator topology on the finite-dimensional state spaces, with the
  same anomaly-line convention on both sides;
\item the cylinder converges to the identity and disjoint union converges to
  tensor product.
\end{enumerate}
Then the limiting amplitudes define an honest or relative Atiyah--Segal TQFT,
according to whether the anomaly line has been trivialized.  Indeed, the
limiting assignments are \(\Sigma\mapsto\mathcal H(\Sigma)\) on objects and
\(M\mapsto Z(M)\) on morphisms; the four hypotheses are precisely the
well-definedness, composition, identity, and symmetric monoidal requirements
in Definition~\ref{def:atiyah-segal-tqft}, with the representative dependence
recorded as the relative anomaly of
Definition~\ref{def:anomalous-relative-tqft} when the anomaly line is kept.

\begin{openproblem}[Functorial extraction from interacting QFT]
\label{op:functorial-extraction-from-qft}
Given a local interacting QFT with a proposed topological sector, construct
the associated ordinary or extended bordism functor directly from its Hilbert
spaces, local algebras, or Euclidean functional integrals.  The construction
should identify the precise completion of state spaces, the gluing topology,
the anomaly line, the boundary category, and the status of gauge fixing or BV
integration cycles.
\end{openproblem}

\section{Four-Dimensional Cohomological Gauge Theory As A Test Case}

Witten--Donaldson theory will be a principal test case for the relation
between cohomological path integrals, functorial TQFT data, and ordinary
four-dimensional gauge theory.  Its ultraviolet presentation is the
topological twist of \(\mathcal N=2\) Yang--Mills theory.  The Donaldson side
is organized by the moduli space of anti-self-dual connections, its
compactification and orientation data, and the observables obtained from the
descent of gauge-invariant polynomials.  To define the invariant one must
state the gauge group and global form, the four-manifold and its orientation
data, the metric dependence and chamber structure, the compactification of
the instanton moduli space, and the treatment of reducible connections.

The Seiberg--Witten explanation compares this ultraviolet cohomological gauge
theory to an Abelian monopole effective theory on the Coulomb branch.  The
physical mechanism is an RG statement: after flowing from the non-Abelian
\(\mathcal N=2\) theory to its low-energy Wilsonian effective action, the
singular loci where monopoles or dyons become light supply the Abelian fields
and monopole equations that enter the low-energy topological path integral.
For the monograph this comparison is not certified by naming the flow.  A
complete argument must construct the Wilsonian coordinate \(u\), the effective
coupling \(\tau(u)\), the duality frame near each singular locus, the
monopole or dyon hypermultiplet, the contact terms and gravitational
couplings, the observable map from Donaldson classes to low-energy
observables, and the wall-crossing or \(u\)-plane terms when they are
present.

\begin{openproblem}[Donaldson--SW RG comparison]
\label{op:donaldson-seiberg-witten-rg-comparison}
Develop the Donaldson/Seiberg--Witten relation as a theorem-level QFT
comparison.  The target is to identify which parts follow from rigorous
differential geometry of instanton and monopole moduli spaces, which parts
follow from finite-dimensional localization or gluing, and which parts depend
on a still-unproved RG theorem connecting the twisted non-Abelian
\(\mathcal N=2\) Yang--Mills theory to the Abelian monopole effective theory.
The desired output is an object-level map between correlation functions or
cohomological observables, with stated metric, chamber, compactification,
contact-term, and anomaly data.
\end{openproblem}
