\chapter{Metric Independence and Cohomological Observables}
\label{ch:tqft-metric-independence}

\ConventionsEuclidean

A topological sector of a quantum field theory is not defined by the absence
of formulas containing a metric.  It is defined by correlation functionals
whose variation under allowed changes of geometric representatives is
controlled and, for the observables under consideration, vanishes in
cohomology.  The most common mechanism is cohomological: an odd symmetry
\(Q\) acts on the regulated fields, sources, and insertions; metric variation
of the action and of the insertions is \(Q\)-exact; and the regulated
expectation functional obeys the corresponding Ward identity.  This chapter
develops that mechanism before any particular example is introduced.

\section{Metric Response as a Renormalized Insertion}

Let \(M\) be a closed oriented smooth \(D\)-manifold and let
\(\mathcal G(M)\) denote the Fréchet manifold of smooth Riemannian metrics on
\(M\).  A metric-dependent QFT presentation assigns, for \(g\in\mathcal G(M)\),
renormalized correlation functionals
\[
  \langle X\rangle_g
\]
for insertions \(X\) in a specified algebra of local, extended, and source
dependent observables.  The notation includes the regulator, local
counterterms, source chart, and contact-term convention.  Different choices
can present the same continuum theory, but a metric-variation formula is a
statement inside one such presentation.

\begin{definition}[Stress-tensor response]
\label{def:stress-tensor-response}
Let \(X\) be a renormalized insertion whose support is disjoint from the
support of a compactly supported metric variation \(\delta g^{\mu\nu}\).
The stress-tensor response of the presentation is the distributional
identity
\begin{equation}
  \delta_g\langle X\rangle_g
  =
  -\frac12
  \int_M\dd^Dx\,\sqrt g\,\delta g^{\mu\nu}(x)\,
  \langle T_{\mu\nu}(x)X\rangle_g
  +
  \langle C_{\delta g}(X)\rangle_g .
  \label{eq:tqft-stress-response}
\end{equation}
Here \(T_{\mu\nu}\) is the renormalized stress tensor and
\(C_{\delta g}(X)\) is the sum of contact insertions supported where the
metric variation collides with the support or source-dependence of \(X\).
\end{definition}

If \(X\) is a product of metric-independent local operators at fixed points
and \(\delta g\) is supported away from those points, then
\(C_{\delta g}(X)=0\).  If \(X\) includes integrated observables, line
operators, boundary conditions, or metric-dependent representatives, the
contact term is part of the definition of the observable.  Ignoring it is
equivalent to changing the problem.

For a one-parameter family \(g_s\), \eqref{eq:tqft-stress-response} becomes
\begin{equation}
  \frac{\dd}{\dd s}\langle X_s\rangle_{g_s}
  =
  -\frac12\int_M\dd^Dx\sqrt{g_s}\,
  \dot g_s^{\mu\nu}\langle T_{\mu\nu}(x)X_s\rangle_{g_s}
  +
  \left\langle C_{\dot g_s}(X_s)+\dot X_s\right\rangle_{g_s}.
  \label{eq:tqft-one-parameter-metric-response}
\end{equation}
The term \(\dot X_s\) records explicit variation of the chosen insertion
representative.  Metric independence is the assertion that the right-hand
side vanishes for the chosen cohomology classes and for every allowed
path \(g_s\).

\section{Cohomological Ward Data}

The letter \(Q\) will denote an odd operator of degree \(+1\).  If \(Y\) is a
homogeneous insertion, write \(|Y|\in\mathbb Z_2\) for its parity.  The
graded commutator with \(Q\) is
\[
  [Q,Y]_{\rm gr}=QY-(-1)^{|Y|}YQ .
\]
For an odd \(Y\), this is an anticommutator; for an even \(Y\), it is a
commutator.  In a path-integral presentation it is often clearer to regard
\(Q\) as an odd derivation \(\delta_Q\) on the algebra of insertions.

\begin{definition}[Cohomological Ward datum]
\label{def:cohomological-ward-datum}
A cohomological Ward datum over a family of metrics \(g_s\) consists of:
\begin{enumerate}[label=(\roman*)]
\item a graded algebra \(\mathcal A_s\) of renormalized insertions;
\item an odd degree-\(+1\) derivation
  \(\delta_Q:\mathcal A_s\to\mathcal A_s\) satisfying
  \(\delta_Q^2=0\);
\item expectation functionals
  \(\langle\cdot\rangle_s:\mathcal A_s\to\mathbb C\);
\item a subspace \(\mathcal A_s^{\rm adm}\subset\mathcal A_s\) of insertions
  for which metric derivatives and \(Q\)-variations are both defined;
\item the Ward identity
  \begin{equation}
    \langle\delta_Q Y\rangle_s=0,
    \qquad
    Y\in\mathcal A_s^{\rm adm}.
    \label{eq:cohomological-ward-identity}
  \end{equation}
\end{enumerate}
The datum is anomaly-free on this family if
\eqref{eq:cohomological-ward-identity} holds without additional local or
boundary terms.  If
\(\langle\delta_QY\rangle_s=\langle\mathcal A_Q(Y)\rangle_s\) for a nonzero
local functional \(\mathcal A_Q\), then \(\mathcal A_Q\) is the
cohomological anomaly in the chosen presentation.
\end{definition}

The Ward identity is a theorem only after regularization and renormalization
are specified.  In finite-dimensional compact models it is Stokes' theorem.
In gauge theory it is a statement about the BV or BRST measure, the quantum
master equation, and boundary behavior at infinity or at singular strata.

\begin{definition}[Cohomological topological sector]
\label{def:cohomological-topological-sector}
A cohomological topological sector for the family \(g_s\) is a
cohomological Ward datum together with, for every \(s\), a class of
\(Q\)-closed insertions \(X_s\) and an odd insertion \(V_s\) such that
\begin{equation}
  \frac{\dd}{\dd s}\langle X_s\rangle_s
  =
  \left\langle
    \delta_Q V_s
  \right\rangle_s .
  \label{eq:cohomological-topological-sector-condition}
\end{equation}
If the metric response is expressed by a stress tensor, this condition is
usually implemented by
\begin{equation}
  -\frac12\sqrt{g_s}\,\dot g_s^{\mu\nu}T_{\mu\nu}\,X_s
  +
  C_{\dot g_s}(X_s)+\dot X_s
  =
  \delta_Q V_s
  \label{eq:stress-response-q-exact-family}
\end{equation}
as an integrated renormalized insertion.
\end{definition}

This definition keeps two distinct statements separate.  The local formula
\(T_{\mu\nu}=[Q,G_{\mu\nu}]_{\rm gr}\) is sufficient only when the contact
terms and the explicit variation of \(X_s\) are also \(Q\)-exact.  In
examples, the latter terms are often where the nontrivial geometry of the
observable enters.

\begin{theorem}[Metric independence from a cohomological Ward datum]
\label{thm:metric-independence-from-ward-datum}
Let \(g_s\), \(0\le s\le1\), be a smooth path of metrics on a closed
manifold \(M\).  Suppose \(X_s\in\mathcal A_s^{\rm adm}\) satisfies
\(\delta_QX_s=0\), the one-parameter metric response is defined for all
\(s\), and the response insertion is \(Q\)-exact as in
\eqref{eq:cohomological-topological-sector-condition}.  If the Ward identity
\eqref{eq:cohomological-ward-identity} holds for the corresponding \(V_s\),
then
\[
  \langle X_1\rangle_{g_1}
  =
  \langle X_0\rangle_{g_0}.
\]
\end{theorem}

\begin{proof}
By hypothesis,
\[
  \frac{\dd}{\dd s}\langle X_s\rangle_s
  =
  \langle\delta_QV_s\rangle_s .
\]
The Ward identity makes the right-hand side vanish pointwise in \(s\).
Therefore the function \(s\mapsto\langle X_s\rangle_s\) is constant on the
interval.  Integrating over \(s\) gives the equality of the endpoint
correlators.
\end{proof}

In local stress-tensor language, Theorem~\ref{thm:metric-independence-from-ward-datum}
is often summarized by saying that \(T_{\mu\nu}\) is \(Q\)-exact.  The
precise statement is \eqref{eq:stress-response-q-exact-family}: the whole
renormalized metric response, including contact terms and moving insertion
representatives, must be \(Q\)-exact.

\section{Contact Terms, Anomalies, and Boundaries of Field Space}

There are three common mechanisms by which the proof above can fail.

First, the \(Q\)-variation of the measure may be anomalous.  In BV language,
the anomaly is the obstruction to the quantum master equation for the
renormalized action.  If \(S\) is a regulated BV action and \(\Delta\) is the
regulated BV Laplacian, the formal Ward identity follows from
\[
  \frac12(S,S)-\hbar\Delta S=0.
\]
Failure of this equation gives a local functional \(\mathcal A_Q\) whose
cohomology class controls whether a local counterterm can restore the Ward
identity.  This is why gauge-theoretic cohomological theories must be
connected to the BV formalism developed in
Chapter~\ref{ch:bv-master-formalism}.

Second, the integration space may have boundaries or noncompact ends.  A
finite-dimensional model with an odd vector field \(Q\) obeys
\[
  \int_{\mathcal F}\delta_Q(\cdots)=
  \int_{\partial\mathcal F}\iota_Q(\cdots)
\]
whenever Stokes' theorem applies.  Thus the Ward identity requires either
\(\partial\mathcal F=\varnothing\), boundary conditions that kill the
boundary term, or compactification data for which the boundary contribution
is itself cancelled by other strata.  In Donaldson--Witten theory this issue
is reflected in compactification of instanton moduli spaces; in
localization integrals it appears as contour, Jeffrey--Kirwan, or other
cycle choices.

Third, contact terms can change the observable algebra.  If a metric
variation moves the representative of an integrated observable, its
collision with other insertions can generate lower-dimensional descendant
operators.  A cohomological topological sector therefore contains not only
the cohomology of isolated operators but also a contact-term prescription
compatible with products and collisions.

\begin{warning}[Metric independence is not automatic functoriality]
\label{warn:metric-independence-not-functoriality}
Metric-independent closed-manifold numbers do not by themselves define a
topological field theory in the Atiyah--Segal sense.  A functorial theory
must also assign state spaces to codimension-one manifolds, linear maps to
bordisms, and gluing isomorphisms.  Cohomological Ward identities are a
mechanism for producing such data under additional compactness, anomaly,
boundary, and finite-dimensionality hypotheses; they are not a substitute
for those hypotheses.
\end{warning}

\section{Descent and Integrated Observables}

Cohomological gauge theories often start from a local \(Q\)-closed scalar
operator and produce observables supported on cycles.  This is the descent
construction.  It is the source of Donaldson observables, topological
sigma-model observables, and many protected line and surface insertions.

\begin{definition}[Descent datum]
\label{def:tqft-descent-datum}
Let \(O^{(p)}\) be a sequence of renormalized \(p\)-form-valued local
operators, \(0\le p\le D\), with total parity including form degree.  A
descent datum is the system
\begin{equation}
  \delta_Q O^{(0)}=0,
  \qquad
  \delta_Q O^{(p+1)}+\dd O^{(p)}=0,
  \quad 0\le p\le D-1,
  \label{eq:tqft-descent-equations}
\end{equation}
as an identity of renormalized local operator-valued forms, including
contact terms when other insertions are present.
\end{definition}

For a closed oriented \(p\)-cycle \(\Sigma_p\subset M\), define
\[
  O(\Sigma_p)=\int_{\Sigma_p}O^{(p)}.
\]
This integral is a renormalized insertion; if \(\Sigma_p\) intersects other
operator supports, one must specify the collision prescription.

\begin{proposition}[Homological invariance of descended observables]
\label{prop:homological-invariance-descended-observables}
Assume \eqref{eq:tqft-descent-equations} holds and that no contact terms are
encountered during the deformations below.  If \(\Sigma_p\) is closed, then
\(O(\Sigma_p)\) is \(Q\)-closed.  If
\(\Sigma_p'-\Sigma_p=\partial B_{p+1}\), then
\[
  O(\Sigma_p')-O(\Sigma_p)
  =
  -\delta_Q\int_{B_{p+1}}O^{(p+1)} .
  \label{eq:tqft-homologous-descendants-q-exact}
\]
Thus the \(Q\)-cohomology class of \(O(\Sigma_p)\) depends only on the
homology class \([\Sigma_p]\).
\end{proposition}

\begin{proof}
For a closed cycle,
\[
  \delta_Q O(\Sigma_p)
  =
  \int_{\Sigma_p}\delta_QO^{(p)}
  =
  -\int_{\Sigma_p}\dd O^{(p-1)}
  =
  -\int_{\partial\Sigma_p}O^{(p-1)}
  =
  0 .
\]
If \(\Sigma_p'-\Sigma_p=\partial B_{p+1}\), then Stokes' theorem and the
descent equation give
\[
  O(\Sigma_p')-O(\Sigma_p)
  =
  \int_{\partial B_{p+1}}O^{(p)}
  =
  \int_{B_{p+1}}\dd O^{(p)}
  =
  -\delta_Q\int_{B_{p+1}}O^{(p+1)} .
\]
The calculation is an identity in the renormalized insertion algebra under
the stated no-contact hypothesis.
\end{proof}

\begin{example}[Donaldson-type descent form]
\label{ex:donaldson-descent-form}
In a four-dimensional cohomological gauge theory obtained from a twisted
\(\mathcal N=2\) vector multiplet, the scalar field \(\phi\) is even and the
connection \(A\) has an odd one-form partner \(\psi\).  With gauge-invariant
quadratic polynomial
\[
  O^{(0)}=\frac{1}{8\pi^2}\operatorname{tr}\phi^2,
\]
and with the convention
\[
  \delta_QA=\psi,\qquad
  \delta_Q\psi=-D_A\phi,\qquad
  \delta_Q\phi=0,
\]
one may choose the first descendants to be
\[
  O^{(1)}=\frac{1}{4\pi^2}\operatorname{tr}(\phi\psi),
  \qquad
  O^{(2)}=\frac{1}{4\pi^2}\operatorname{tr}
  \left(\phi F+\frac12\psi\wedge\psi\right),
\]
with higher descendants determined by
\eqref{eq:tqft-descent-equations}.  Direct substitution gives
\[
  \delta_QO^{(1)}=-\dd O^{(0)},
  \qquad
  \delta_QO^{(2)}=-\dd O^{(1)},
\]
using the Bianchi identity \(D_AF=0\) and cyclicity of the trace.  The
normalizations depend on the trace convention and must match the
instanton-number normalization of the gauge theory.  The integral
\(\int_{\Sigma_2}O^{(2)}\) is the basic surface observable entering
Donaldson--Witten theory;
Proposition~\ref{prop:homological-invariance-descended-observables}
explains why its cohomology class depends on the homology class of
\(\Sigma_2\), before compactification and contact-term issues are addressed.
\end{example}

\section{A Finite-Dimensional Model of the Ward Identity}

The following model is not a field theory, but it is the algebraic core of
cohomological localization.  Let \(x^i\) be even coordinates on a compact
oriented finite-dimensional manifold \(X\), and let \(\psi^i\) be odd
coordinates corresponding to \(\dd x^i\).  The de Rham differential is the
odd derivation
\[
  Qx^i=\psi^i,\qquad Q\psi^i=0.
\]
An ordinary differential form \(\alpha\) is represented as a function
\(\alpha(x,\psi)\), and \(Q\alpha\) is \(\dd\alpha\).  If \(X\) is closed,
Stokes' theorem gives
\[
  \int_X Q\eta=\int_X\dd\eta=0.
\]
If \(\alpha\) is \(Q\)-closed and \(V_t\) is an odd family, then
\[
  \frac{\dd}{\dd t}
  \int_X \alpha\,\exp[-Q V_t]
  =
  -\int_X \alpha\,Q(\dot V_t)\,\exp[-QV_t]
  =
  -\int_X Q\left(\alpha\,\dot V_t\,\exp[-QV_t]\right)=0,
\]
where \(Q\alpha=0\), \(Q(QV_t)=0\), and the graded Leibniz rule fix the
sign.  This is the finite-dimensional prototype of
Theorem~\ref{thm:metric-independence-from-ward-datum}.  Infinite-dimensional
QFT applications require a regulator, a measure or BV half-density, and a
proof that no boundary terms appear from noncompact field directions.

\section{Functorial Target}

The Atiyah--Segal form of a topological field theory assigns a vector space
\(Z(\Sigma)\) to a closed \((D-1)\)-manifold \(\Sigma\) and a linear map
\[
  Z(M):Z(\Sigma_{\rm in})\to Z(\Sigma_{\rm out})
\]
to a bordism \(M:\Sigma_{\rm in}\to\Sigma_{\rm out}\), with compatibility
under gluing:
\[
  Z(M_2\circ M_1)=Z(M_2)Z(M_1).
\]
Extended versions refine this assignment to manifolds with corners and
higher-categorical targets.  The cohomological construction above is one
source of such functors after gauge fixing, anomaly cancellation, boundary
conditions, compactness, and finite-dimensionality hypotheses have been
supplied.

\begin{openproblem}[From cohomological path integrals to extended functors]
\label{op:cohomological-to-extended-functor}
Given a cohomological gauge theory with a \(Q\)-exact metric response,
determine the exact hypotheses under which its correlation functions define
an extended topological functor: compactness or properness of the moduli
problem, anomaly cancellation, boundary conditions, gluing analysis,
orientation data, contact-term compatibility, and compatibility with local
operator insertions.  The objective is not merely to match known invariants
but to construct the functorial QFT object from regulated cohomological
field theory data.
\end{openproblem}
