\chapter{Boundaries, Defects, And Categories In Topological QFT}
\label{ch:tqft-boundaries-defects-categories}

\ConventionsEuclidean

This chapter develops boundary and defect data in topological QFT through
finite-dimensional models where the algebra can be written explicitly.  The
central mechanism is that cutting a topological theory along strata produces
state spaces, module categories, defect bimodules, junction spaces, and
gluing pairings, and these structures must satisfy concrete associativity and
nondegeneracy identities.

\section{Boundary Conditions As Modules}

Let \(Z\) be an extended \(n\)-dimensional topological theory with a chosen
target \((\infty,n)\)-category.  A boundary condition \(B\) for \(Z\) assigns
compatible data to manifolds with boundary labelled by \(B\).  In a fully
local formulation, if the value of \(Z\) on a point is an algebraic object
\(\mathcal A\), then a boundary condition is represented by a module object
over \(\mathcal A\), in the target appropriate to the codimension.

\begin{definition}[Boundary category in dimension one]
\label{def:tqft-boundary-category}
Let \(Z\) be a one-dimensional topological theory with a set of boundary
labels \(\mathcal B\).  For \(B_0,B_1\in\mathcal B\), let
\[
  \mathcal H(B_0,B_1).
\]
be the state space assigned to an interval whose incoming endpoint is
labelled by \(B_0\) and outgoing endpoint by \(B_1\).  Gluing intervals gives
linear maps
\[
  \mathcal H(B_0,B_1)\otimes \mathcal H(B_1,B_2)
  \longrightarrow
  \mathcal H(B_0,B_2).
\]
If these maps are associative and if the interval of zero length gives
identity elements in \(\mathcal H(B,B)\), then the labels \(B\) are the
objects of a category and \(\mathcal H(B_0,B_1)\) are its morphism spaces.
\end{definition}

\begin{remark}[Associativity from one-dimensional gluing]
\label{rem:interval-gluing-category}
The interval amplitudes of a one-dimensional topological theory with labelled
endpoints satisfy the associativity condition in
Definition~\ref{def:tqft-boundary-category}.

Let \(I=I_1\cup I_2\cup I_3\) be an interval decomposed into three subintervals
with intermediate labels \(B_1,B_2\).  Topological invariance says that the
linear map assigned to \(I\) depends only on the labelled bordism, not on the
choice of cut positions.  The two parenthesizations
\[
  (I_1\cup I_2)\cup I_3,
  \qquad
  I_1\cup(I_2\cup I_3)
\]
are the same labelled bordism.  Functoriality therefore identifies the two
composite multiplication maps from
\(\mathcal H(B_0,B_1)\otimes\mathcal H(B_1,B_2)\otimes
\mathcal H(B_2,B_3)\) to \(\mathcal H(B_0,B_3)\).  The zero-length interval
is the identity bordism, hence gives identity morphisms.
\end{remark}

\section{\texorpdfstring{\(BF_2\)}{BF2} and Finite Gauge Boundary States}

The most concrete two-dimensional example is untwisted finite gauge theory,
the finite counterpart of \(BF_2\).  Let \(G\) be a finite group.  The closed
state space on the circle is the center of the group algebra
\[
  \mathcal H(S^1)=Z(\C[G]).
\]
The pair-of-pants product is multiplication in \(Z(\C[G])\).  For each
irreducible complex representation \(\rho\) of \(G\), with character
\(\chi_\rho\) and dimension \(d_\rho\), define
\begin{equation}
  e_\rho
  =
  \frac{d_\rho}{|G|}
  \sum_{g\in G}\chi_\rho(g^{-1})\,g
  \in Z(\C[G]).
  \label{eq:finite-bf-two-idempotent}
\end{equation}

The elements \(e_\rho\), as \(\rho\) ranges over irreducible representations
of \(G\), are the primitive central idempotents of \(\C[G]\):
\[
  e_\rho e_\sigma=\delta_{\rho\sigma}e_\rho,
  \qquad
  \sum_{\rho\in\operatorname{Irr}(G)}e_\rho=1 .
\]
The elementary closed-channel topological boundary states of this semisimple
model are therefore labelled by \(\operatorname{Irr}(G)\).  Equivalently, the
elementary modules over the closed Frobenius algebra \(Z(\C[G])\) are
\[
  M_\rho=e_\rho Z(\C[G]),
  \qquad
  \operatorname{Hom}_{Z(\C[G])}(M_\rho,M_\sigma)
  \cong
  \begin{cases}
  \C, & \rho=\sigma,\\
  0, & \rho\ne\sigma .
  \end{cases}
\]
The regular representation decomposes as
\[
  \C[G]\cong \bigoplus_{\rho\in\operatorname{Irr}(G)}
  V_\rho\otimes V_\rho^* .
\]
The center acts on the summand \(V_\rho\otimes V_\rho^*\) by a scalar.
Schur orthogonality gives the central projector onto that summand as
\eqref{eq:finite-bf-two-idempotent}.  Projectors onto distinct summands are
orthogonal and their sum is the identity.  Since \(Z(\C[G])\) is a finite
direct product of copies of \(\C\), its simple modules are exactly the
one-dimensional summands \(e_\rho Z(\C[G])\), with the displayed Hom spaces.

This calculation is deliberately a closed-sector statement.  A fully extended
finite-gauge theory has a richer boundary classification by module categories
over \(\operatorname{Rep}(G)\), and twisted theories replace
\(\operatorname{Rep}(G)\) by the corresponding cocycle-twisted tensor
category.  The idempotents above are the part of that story visible to the
ordinary two-dimensional closed TQFT.

\section{Defect Fusion and Junction Spaces}

A codimension-one defect \(D_{12}\) from a theory \(Z_1\) to a theory \(Z_2\)
is placed on a separating hypersurface.  Fusion is defined by bringing two
parallel defects together:
\[
  D_{12}\circ D_{23}=D_{13}.
\]
For a topological defect, correlation functions are unchanged under
deformations that keep the ordering of defects and avoid operator insertions.
The fusion product is associative when the three-defect collision limit is
independent of the relative distances used to take the limit.  In a
regulated QFT this independence is a theorem to be proved; in a bordism
presentation it is part of the functorial data.

\begin{example}[Finite gauge defects as bisets]
\label{ex:finite-gauge-biset-defects}
Let \(G\) and \(H\) be finite groups.  A topological interface from
two-dimensional finite \(G\)-gauge theory to finite \(H\)-gauge theory can be
modelled by a finite \((G,H)\)-biset \(X\), namely a finite set with commuting
left \(G\)- and right \(H\)-actions.  If \(Y\) is an \((H,K)\)-biset, the
fused interface is the balanced product
\[
  X\otimes_H Y
  =
  (X\times Y)/\{(xh,y)\sim(x,hy):h\in H\},
\]
which is a \((G,K)\)-biset.  Junction fields between two \((G,H)\)-bisets
\(X\) and \(X'\) are the equivariant functions
\[
  \operatorname{Hom}_{G,H}(X,X')
  =
  \{f:X\to X': f(gxh)=gf(x)h\}.
\]
\end{example}

\paragraph{Associativity of finite defect fusion.}
For finite biset defects the balanced-product convention gives
\[
  (X\otimes_H Y)\otimes_K Z
  \cong
  X\otimes_H (Y\otimes_K Z)
\]
canonically as \((G,L)\)-bisets, whenever \(X,Y,Z\) are composable
\((G,H)\)-, \((H,K)\)-, and \((K,L)\)-bisets.
Both sides are quotients of \(X\times Y\times Z\) by the equivalence relation
generated by
\[
  (xh,y,z)\sim(x,hy,z),
  \qquad
  (x,yk,z)\sim(x,y,kz),
  \qquad h\in H,\ k\in K .
\]
The identity map on \(X\times Y\times Z\) therefore descends to a bijection
between the two iterated quotients.  The left \(G\)- and right \(L\)-actions
commute with the two generating equivalences, so the bijection is a
\((G,L)\)-biset isomorphism.  This is quotient bookkeeping, not a theorem
about the existence of defect fusion in an arbitrary continuum QFT; the latter
requires the topological-defect sector and its fusion operation to have been
constructed independently.

\section{Line Defects and Modular Tensor Data}

In a three-dimensional oriented TQFT, line defects form a monoidal category
\(\mathcal C\).  The tensor product is parallel fusion:
\[
  X\otimes Y=\lim_{\epsilon\to0}X(\epsilon)Y(0).
\]
Braiding is obtained by exchanging two topological lines in \(\R^3\):
\[
  c_{X,Y}:X\otimes Y\longrightarrow Y\otimes X.
\]
The hexagon identities are equalities of isotopies of three braided lines.
If every line has a dual line and the vacuum line has endomorphism ring
\(\C\), then \(\mathcal C\) is rigid and has a distinguished unit object.
Additional finiteness, semisimplicity, and nondegenerate braiding hypotheses
give a modular tensor category; these hypotheses are part of the claimed
three-dimensional theory, not consequences of topology alone.

\begin{definition}[Modular tensor category data]
\label{def:modular-tensor-category-data}
For the purposes of this volume, a semisimple modular tensor category
\(\mathcal C\) consists of a finite set \(I\) of simple objects, a tensor unit
\(0\in I\), duals \(a\mapsto a^*\), fusion multiplicities
\(N_{ab}^{\phantom{ab}c}\), associativity isomorphisms \(F\), braiding
isomorphisms \(R\), twists \(\theta_a\), and a nondegenerate matrix
\[
  S_{ab}=\frac{1}{\mathcal D}
  \operatorname{tr}_{a\otimes b}\!\left(c_{b,a}c_{a,b}\right),
  \qquad
  \mathcal D^2=\sum_{a\in I}d_a^2,
\]
subject to the pentagon, hexagon, ribbon, and nondegeneracy identities.  The
numbers \(d_a\) are categorical dimensions.  A three-dimensional TQFT whose
line category is \(\mathcal C\) must produce these data from line fusion,
line exchange, and framed line rotation.
\end{definition}

\begin{example}[\(SU(2)_k\) line data]
\label{ex:su2k-line-data-boundary-chapter}
For \(SU(2)\) Chern--Simons theory at positive integer level \(k\), the line
labels are \(a=0,\ldots,k\), corresponding to spin \(a/2\).  The modular
matrix is
\[
  S_{ab}
  =
  \sqrt{\frac{2}{k+2}}\,
  \sin\!\left(\frac{(a+1)(b+1)\pi}{k+2}\right),
\]
and the fusion rule is
\[
  N_{ab}^{\phantom{ab}c}=1
  \quad\Longleftrightarrow\quad
  |a-b|\le c\le \min(a+b,2k-a-b),
  \quad a+b+c\in2\mathbb Z,
\]
with \(N_{ab}^{\phantom{ab}c}=0\) otherwise.  The derivation and the Hopf-link
normalization are given in
Chapter~\ref{ch:tqft-chern-simons-theory}; here the same data are read as the
input controlling boundaries, condensable algebras, and line-defect junctions.
\end{example}

\section{Boundary Categories for Chern--Simons Theory}

For compact \(G\) Chern--Simons theory at level \(k\), a Wilson line labelled
by a representation \(R\) is
\[
  W_R(\gamma)=\operatorname{tr}_R\,\mathcal P
  \exp\!\left(\int_\gamma A\right).
\]
Quantum gauge invariance restricts the admissible line labels.  For simply
connected compact simple \(G\), the expected finite label set is the set of
integrable highest weights at level \(k\).  Fusion of line defects is then
encoded by the corresponding representation category of the affine algebra or
quantum group at a root of unity, depending on the construction chosen.

The boundary chiral algebra is not an extra decoration.  It is the algebraic
object that describes states on a surface with boundary and line endpoints.
The equality between bulk Chern--Simons amplitudes and boundary conformal
blocks is a gluing statement relating two constructions of the same
surface-state spaces.

\begin{definition}[Lagrangian subgroup of an abelian Chern--Simons theory]
\label{def:abelian-cs-lagrangian-subgroup}
Let \(K\) be a nondegenerate integral symmetric \(r\times r\) matrix defining
an abelian Chern--Simons theory.  The finite group of line labels is
\[
  A_K=\mathbb Z^r/K\mathbb Z^r .
\]
For charge representatives \(\ell,\ell'\in\mathbb Z^r\), define
\[
  b(\ell,\ell')=\ell^T K^{-1}\ell'\quad\operatorname{mod}\mathbb Z,
  \qquad
  q(\ell)=\frac12\ell^T K^{-1}\ell\quad\operatorname{mod}\mathbb Z .
\]
A subgroup \(L\subset A_K\) is Lagrangian if \(q|_L=0\) and
\[
  L=L^\perp,
  \qquad
  L^\perp=\{a\in A_K: b(a,\ell)=0\ \text{for every}\ \ell\in L\}.
\]
\end{definition}

\paragraph{Abelian Chern--Simons boundary algebra.}
In the pointed modular category of an abelian Chern--Simons theory, a
Lagrangian subgroup \(L\subset A_K\) determines a commutative separable
algebra object
\[
  A_L=\bigoplus_{\ell\in L}\ell .
\]
The condition \(q|_L=0\) is exactly the condition that the summands have
trivial topological spin, and \(L=L^\perp\) is the condition that no nontrivial
bulk line has trivial monodromy with every condensed line.  This follows from
the finite semisimple algebra of the pointed category: the simple objects form
an abelian group under
fusion, here \(A_K\).  The braiding monodromy of charges \(\ell,\ell'\) is
\(\exp(2\pi\ii b(\ell,\ell'))\), and the twist of \(\ell\) is
\(\exp(2\pi\ii q(\ell))\).  The direct sum over a subgroup carries
multiplication induced by the group law \(L\times L\to L\).  Commutativity of
this multiplication requires trivial mutual braiding among condensed
summands; this follows from \(q|_L=0\) by the quadratic identity
\[
  b(\ell,\ell')=q(\ell+\ell')-q(\ell)-q(\ell') .
\]
Separability is the finite semisimple averaging idempotent for the finite
group \(L\).  A line \(a\) can pass through the boundary condensate without
monodromy precisely when \(a\in L^\perp\).  If \(L=L^\perp\), the only such
transparent bulk lines are the condensed ones, so the boundary is a boundary
to the trivial bulk rather than to another residual topological order.

\begin{example}[Toric-code boundary subgroups]
\label{ex:toric-code-lagrangian-boundaries}
For
\[
  K=\begin{pmatrix}0&2\\2&0\end{pmatrix},
\]
the line group is \(A_K\cong\mathbb Z_2\oplus\mathbb Z_2\).  Writing the
generators as \(e=(1,0)\) and \(m=(0,1)\), one has
\[
  q(e)=q(m)=0,
  \qquad
  b(e,m)=\frac12 .
\]
Thus
\[
  L_e=\{0,e\},
  \qquad
  L_m=\{0,m\}
\]
are Lagrangian subgroups, while \(\{0,e+m\}\) is not because
\(q(e+m)=1/2\).  The two Lagrangian subgroups are the two elementary gapped
boundaries: one condenses the electric line and the other condenses the
magnetic line.
\end{example}

\section{BV-BFV Boundary Mechanism}

For a gauge theory on a manifold \(M\) with boundary, the BV variation has the
form
\[
  \delta S_{\mathrm{BV}}
  =
  \iota_{Q_{\mathrm{BV}}}\omega_{\mathrm{BV}}
  +
  \pi_{\partial M}^*\alpha_{\partial}.
\]
The boundary one-form \(\alpha_{\partial}\) has differential
\[
  \omega_{\partial}=\delta\alpha_{\partial}.
\]
The pair \((\mathcal F_{\partial},\omega_{\partial})\) is the BFV phase
space, and the bulk BV vector field induces a boundary cohomological vector
field when the master equation is compatible with the boundary term.  A
boundary condition is a Lagrangian subspace or derived Lagrangian object
\[
  \mathcal L_{\partial}\subset\mathcal F_{\partial}
\]
on which the boundary variation vanishes.  This is the local mechanism behind
boundary categories in gauge-theoretic TQFTs.

\section{Crane--Yetter and Walker--Wang Boundaries}

Four-dimensional state-sum and lattice-Hamiltonian constructions make the
boundary role of braided tensor categories especially visible.  The fully
general Crane--Yetter state sum and Walker--Wang Hamiltonian construction
belong to a broad categorical theorem package.  The local mechanism needed
for the present monograph can be proved directly in the pointed case, which
is already rich enough to distinguish bulk deconfinement from boundary
topological order.

\begin{definition}[Pointed braided input and its radical]
\label{def:pointed-braided-radical}
Let \(A\) be a finite abelian group.  A pointed braided input is a quadratic
function
\[
  q:A\to\mathbb Q/\mathbb Z
\]
whose polarization
\begin{equation}
  b(a,x)=q(a+x)-q(a)-q(x)
  \label{eq:pointed-braiding-polarization}
\end{equation}
is a symmetric biadditive pairing \(A\times A\to\mathbb Q/\mathbb Z\).  The
topological spin of the simple line \(a\) is \(\theta_a=\exp(2\pi\ii q(a))\),
and the double braiding of \(a\) around \(x\) is
\[
  M(a,x)=\exp(2\pi\ii b(a,x)).
\]
The radical, or transparent subgroup, is
\begin{equation}
  \operatorname{Rad}(A,b)
  =
  \{a\in A:\ b(a,x)=0\ {\rm in}\ \mathbb Q/\mathbb Z
  \ {\rm for\ every}\ x\in A\}.
  \label{eq:pointed-radical}
\end{equation}
The pointed input is nondegenerate when \(\operatorname{Rad}(A,b)=0\).
\end{definition}

\begin{construction}[Pointed Walker--Wang boundary mechanism]
\label{cons:pointed-walker-wang-boundary-mechanism}
Fix the pointed braided finite input of
Definition~\ref{def:pointed-braided-radical}.  The finite Walker--Wang
mechanism used here is the following algebraic lattice model.  Oriented edge
labels take values in \(A\).  Vertex terms enforce the group-law fusion
constraint, and a plaquette move inserts a closed loop of label \(x\in A\) and
recouples it through the existing labels.  When a candidate line of label
\(a\) crosses that loop, the local move multiplies the state by the
double-braiding phase
\begin{equation}
  M(a,x)=\exp(2\pi\ii b(a,x)).
  \label{eq:pointed-ww-crossing-phase}
\end{equation}
In this finite mechanism a bulk line label \(a\) is called deconfined when its
line insertion commutes with every plaquette recoupling away from its endpoints;
equivalently, translating a segment of the line through the bulk leaves no
string of plaquette violations.  The construction therefore gives the finite
criterion
\begin{equation}
  a\hbox{ is deconfined}
  \quad\Longleftrightarrow\quad
  M(a,x)=1\quad\hbox{for every }x\in A
  \quad\Longleftrightarrow\quad
  a\in\operatorname{Rad}(A,b).
  \label{eq:pointed-ww-radical-criterion}
\end{equation}
If \(b\) is nondegenerate, this finite algebraic model has no nontrivial
deconfined bulk line.  On a boundary where strings are allowed to end, the
endpoint labels still carry the pointed braided category \((A,q)\): fusion is
addition in \(A\), spins are \(\exp(2\pi\ii q(a))\), and double braiding is
\(\exp(2\pi\ii b(a,x))\).
\end{construction}

\paragraph{Verification of the finite criterion.}
The vertex constraint only uses the group law of \(A\), so a closed plaquette
loop can be slid through a candidate line one elementary crossing at a time.
Each crossing contributes the phase
\eqref{eq:pointed-ww-crossing-phase}.  If \(M(a,x)\ne1\) for some \(x\), the
plaquette term with loop label \(x\) detects the line segment; moving the
segment changes the local plaquette eigenvalue along its path, which is the
finite-lattice meaning of confinement in this mechanism.  If \(M(a,x)=1\) for
all \(x\), all plaquette recouplings away from the endpoints commute with the
line insertion, so the energy cost is localized at the endpoints and the line
is deconfined.  The equality \(M(a,x)=1\) for every \(x\) is precisely
\(b(a,x)=0\) in \(\mathbb Q/\mathbb Z\) for every \(x\), hence
\eqref{eq:pointed-ww-radical-criterion}.  A boundary removes the plaquette
relation that would continue the detection surface outside the material, so a
string may terminate there; the same local crossing calculation records the
boundary spin and double-braiding data.  Thus the finite mechanism separates a
radical bulk spectrum from a boundary pointed braided category.

\begin{example}[Toric-code input as a modular boundary order]
\label{ex:toric-code-walker-wang-boundary}
For \(A=\mathbb Z_2e\oplus\mathbb Z_2m\), set
\[
  q(e)=q(m)=0,\qquad q(e+m)=\frac12 .
\]
Then \(b(e,m)=1/2\), while \(b(e,e)=b(m,m)=0\).  The radical is zero:
\[
  \operatorname{Rad}(A,b)=\{0\}.
\]
The pointed Walker--Wang mechanism therefore has no nontrivial deconfined
bulk line, but the boundary still supports the toric-code anyons
\(\{1,e,m,e+m\}\) with the displayed mutual semionic braiding.
\end{example}

\begin{definition}[M\"uger center]
\label{def:muger-center-braided-input}
Let \(\mathcal C\) be a finite semisimple braided tensor category over
\(\C\), with braiding
\[
  c_{a,x}:a\otimes x\longrightarrow x\otimes a .
\]
For simple objects \(a,x\), the monodromy is the natural automorphism
\[
  M_{a,x}=c_{x,a}c_{a,x}:a\otimes x\longrightarrow a\otimes x .
\]
The M\"uger center \(\mathcal Z_2(\mathcal C)\) is the full fusion
subcategory generated by the simple objects \(a\) such that
\begin{equation}
  M_{a,x}=\operatorname{id}_{a\otimes x}
  \qquad\text{for every simple }x\in\mathcal C .
  \label{eq:muger-center-definition}
\end{equation}
The braided category is nondegenerate when
\(\mathcal Z_2(\mathcal C)\) is generated only by the tensor unit.
\end{definition}

\begin{proposition}[M\"uger-center plaquette criterion]
\label{prop:muger-center-plaquette-criterion}
Assume a Walker--Wang local plaquette algebra whose simple string labels are
the simple objects of a finite semisimple braided category \(\mathcal C\),
and assume that dragging a plaquette loop labelled by \(x\) through a
transverse simple line \(a\) acts on \(a\otimes x\) by the categorical
monodromy \(M_{a,x}=c_{x,a}c_{a,x}\).  Then the line \(a\) commutes with all
local plaquette-loop recouplings exactly when
\[
  a\in\mathcal Z_2(\mathcal C).
\]
Thus the algebraic candidate for a deconfined bulk line is a simple object in
the M\"uger center.  If \(\mathcal C\) is nondegenerate, this local criterion
leaves only the tensor unit as a bulk deconfined line.
\end{proposition}

\begin{proof}
Let \(P_x\) denote the elementary plaquette recoupling in which a small loop
labelled by \(x\) is created, slid across the candidate line \(a\), and
absorbed.  By the stated hypothesis, the part of this operation detected by
the crossing with \(a\) is the monodromy \(M_{a,x}\).  If \(a\) commutes with
every plaquette recoupling, then \(P_x\) acts on any local configuration
containing \(a\) exactly as it acts when the loop is not linked with \(a\).
Therefore the monodromy on \(a\otimes x\) must be the identity for every
simple \(x\), which is precisely
\eqref{eq:muger-center-definition}.

Conversely, if \(a\) satisfies \eqref{eq:muger-center-definition}, then
sliding any simple plaquette loop past \(a\) has trivial monodromy.  A
general plaquette loop is a finite direct sum of simple labels, and the
recoupling maps are natural with respect to this direct-sum decomposition.
The commutator with \(a\) is therefore zero on each simple summand and hence
zero on the whole local plaquette algebra.  Nondegeneracy of
\(\mathcal C\) is exactly the assertion that the only simple object with
trivial monodromy against all simples is the tensor unit.
\end{proof}

\begin{example}[Non-pointed check: the Ising braided category]
\label{ex:ising-muger-center-check}
The Ising modular category has simple labels
\[
  1,\qquad \psi,\qquad \sigma,
\]
with quantum dimensions \(1,1,\sqrt2\) and total dimension
\(\mathcal D=2\).  In the ordered basis \((1,\psi,\sigma)\), its modular
matrix is
\[
  S=\frac12
  \begin{pmatrix}
    1&1&\sqrt2\\
    1&1&-\sqrt2\\
    \sqrt2&-\sqrt2&0
  \end{pmatrix}.
\]
For a transparent simple object \(a\) in a modular category, the Hopf-link
amplitude factorizes as
\begin{equation}
  S_{ax}=\frac{d_a d_x}{\mathcal D}
  \qquad\text{for every simple }x .
  \label{eq:transparent-s-row-factorization}
\end{equation}
The unit row satisfies this identity.  The \(\psi\) row fails at
\(x=\sigma\), because
\[
  S_{\psi\sigma}=-\frac{\sqrt2}{2}
  \ne
  \frac{d_\psi d_\sigma}{\mathcal D}
  =
  \frac{\sqrt2}{2}.
\]
The \(\sigma\) row fails already at \(x=\psi\).  Hence the M\"uger center is
trivial.  The Ising input is non-pointed and nondegenerate, so the local
Walker--Wang plaquette criterion leaves no nontrivial bulk deconfined line,
while the non-pointed Ising anyons may still appear as boundary line data
when the boundary construction is supplied.
\end{example}

\begin{remark}[General categorical status]
\label{rem:crane-yetter-walker-wang-general-status}
For a general spherical, ribbon, or modular fusion category \(\mathcal C\),
the Crane--Yetter and Walker--Wang constructions require the full categorical
state-sum or lattice-Hamiltonian theorem: pentagon and hexagon identities,
Pachner-move invariance, spherical trace normalization, and the analysis of
bulk excitation sectors.  The monograph uses the general theorem as external
mathematics until those categorical proofs are developed in full.  The
pointed theorem above proves the mechanism that is used conceptually here:
bulk deconfined lines are controlled by the M\"uger center, and a nondegenerate
braided input leaves its nontrivial anyons as boundary, not bulk, topological
order.
\end{remark}

\section{Extended Assignment}

An extended \(n\)-dimensional TQFT assigns:
\[
  Z(M^n)\in\C,\qquad
  Z(\Sigma^{n-1})\in\mathrm{Vect},
\]
\[
  Z(Y^{n-2})\in\mathrm{Cat},
\]
and higher categorical objects to lower-dimensional strata, in a target
\((\infty,n)\)-category chosen before the assignment is made.  Dualizability
conditions ensure that gluing along all strata is represented by composition
and evaluation maps.  In a QFT realization, the target objects must be
constructed from regulated state spaces, boundary operator algebras, and
defect junction spaces.

\begin{openproblem}[Gauge-theoretic extended construction]
\label{op:gauge-theoretic-extended-tqft-construction}
Construct the extended boundary and defect categories of four-dimensional
cohomological gauge theories directly from regulated BV-BFV data, including
instanton and monopole boundary strata, line and surface defects, gluing
maps, orientations, and anomaly lines.
\end{openproblem}
