\chapter{Finite Gauge Theory and State-Sum TQFT}
\label{ch:tqft-finite-gauge-state-sum}

\ConventionsEuclidean

Finite gauge theory is the cleanest nonperturbative laboratory for
topological path integrals.  The field spaces are finite groupoids, the
``measure'' is groupoid cardinality, the state spaces are finite-dimensional
vector spaces of functions or sections over boundary groupoids, and gluing is
Fubini's theorem for finite homotopy fibers.  This chapter develops the
untwisted theory and its Dijkgraaf--Witten cocycle twists with enough detail
that later references to finite topological gauge theory have a precise
meaning.

Throughout this chapter \(G\) is a finite group.  A principal \(G\)-bundle
means a right principal bundle.  All manifolds are compact, smooth,
oriented when orientation is needed, and equipped with collars at boundary
components.  If \(X\) is a finite groupoid, \(\pi_0X\) denotes the set of
isomorphism classes of objects and \(\operatorname{Aut}_X(x)\) denotes the
automorphism group of an object \(x\).

\begin{definition}[Finite gauge state-sum datum]
\label{def:finite-gauge-state-sum-datum}
Fix a dimension \(D\).  A finite \(G\)-gauge state-sum datum is a tuple
\[
  \mathfrak Z_{G,\omega}^{(D)}
  =
  \bigl(
    G,[\omega],
    M\mapsto\operatorname{Bun}_G(M),
    \mathcal L_\omega,
    \mathcal H_{G,\omega},
    Z_{G,\omega},
    K\mapsto\operatorname{Flat}_G(K)
  \bigr)
\]
with the following entries.
\begin{enumerate}[label=\textnormal{(\roman*)}]
\item \(G\) is a finite group, and
  \([\omega]\in H^D(BG,U(1))\) is the Dijkgraaf--Witten twist.  The
  untwisted theory is the case \([\omega]=0\).
\item For every compact collared \(D\)-manifold \(M\) and closed
  \((D-1)\)-manifold \(\Sigma\),
  \(\operatorname{Bun}_G(M)\) and \(\operatorname{Bun}_G(\Sigma)\) are the
  finite groupoids of right principal \(G\)-bundles and bundle isomorphisms.
\item \(\mathcal L_\omega(\Sigma)\) is the transgressed boundary line over
  \(\operatorname{Bun}_G(\Sigma)\).  In the untwisted theory this line is
  canonically trivial.
\item The boundary state space is
  \[
    \mathcal H_{G,\omega}(\Sigma)
    =
    \Gamma\bigl(\operatorname{Bun}_G(\Sigma),\mathcal L_\omega(\Sigma)\bigr),
  \]
  with the groupoid-cardinality pairing.  In the untwisted case this is the
  vector space of functions on \(\pi_0\operatorname{Bun}_G(\Sigma)\).
\item A bordism \(M:\Sigma_0\to\Sigma_1\) defines the restriction span
  \[
    \operatorname{Bun}_G(\Sigma_0)
    \xleftarrow{\ r_0\ }
    \operatorname{Bun}_G(M)
    \xrightarrow{\ r_1\ }
    \operatorname{Bun}_G(\Sigma_1),
  \]
  and \(Z_{G,\omega}(M)\) is the corresponding push-pull operator, weighted
  by the cocycle phase and by automorphism denominators.
\item For a branched triangulation \(K\) of \(M\),
  \(\operatorname{Flat}_G(K)\) is the finite set of flat edge labelings whose
  action groupoid \(\operatorname{Flat}_G(K)//G^{K_0}\) presents
  \(\operatorname{Bun}_G(M)\).  The simplicial state sum is a representative
  of the intrinsic groupoid integral.
\end{enumerate}
The gluing theorem below proves that these entries define a symmetric
monoidal TQFT on the corresponding collared bordism category.
\end{definition}

\section{Finite Groupoids and the Gauge Volume}

\begin{definition}[Finite groupoid cardinality]
\label{def:finite-groupoid-cardinality}
A finite groupoid is a groupoid with finitely many isomorphism classes and
finite automorphism groups.  Its cardinality is
\begin{equation}
  |X|
  =
  \sum_{[x]\in\pi_0X}
  \frac{1}{|\operatorname{Aut}_X(x)|}.
  \label{eq:finite-groupoid-cardinality}
\end{equation}
If \(F:X\to\C\) is a function constant on isomorphism classes, its integral
over \(X\) is
\begin{equation}
  \int_X F
  =
  \sum_{[x]\in\pi_0X}
  \frac{F(x)}{|\operatorname{Aut}_X(x)|}.
  \label{eq:finite-groupoid-integral}
\end{equation}
\end{definition}

The factor \(1/|\operatorname{Aut}(x)|\) is the finite gauge-volume factor.
It is forced by gluing.  Omitting it gives the wrong cylinder and the wrong
answer for a quotient with stabilizers.

\paragraph{Action groupoid cardinality.}
Let a finite group \(H\) act on a finite set \(X\).  The action groupoid
\(X//H\) has objects \(X\) and morphisms \(x\to h x\) labeled by
\(h\in H\).  Its groupoid cardinality is
\begin{equation}
  |X//H|=\frac{|X|}{|H|}.
  \label{eq:action-groupoid-cardinality}
\end{equation}

The isomorphism classes are the orbits \(\mathcal O\subset X\).  For
\(x\in\mathcal O\), the automorphism group is the stabilizer \(H_x\).  Thus
\[
  |X//H|
  =
  \sum_{\mathcal O}
  \frac{1}{|H_x|}
  =
  \sum_{\mathcal O}
  \frac{|\mathcal O|}{|H|}
  =
  \frac{|X|}{|H|},
\]
where the orbit-stabilizer theorem was used in the second equality.
This calculation is the finite model for dividing by gauge volume: summing
one over each object and then weighting by automorphisms gives the same result
as averaging over the acting group before passing to orbits.

\section{Fields as Principal-Bundle Groupoids}

For the datum \(\mathfrak Z_{G,\omega}^{(D)}\), and in particular for the
untwisted theory used first, define for a compact manifold \(M\)
\[
  \operatorname{Bun}_G(M)
\]
to be the groupoid of principal \(G\)-bundles \(P\to M\) and bundle
isomorphisms.  Equivalently,
\(\operatorname{Bun}_G(M)\) is the mapping groupoid from \(M\) to the
classifying stack \(BG\).  In this chapter this phrase has the concrete
meaning supplied by principal bundles and their isomorphisms; no analytic
mapping-space structure is being assumed.

\begin{definition}[Untwisted finite gauge partition function]
\label{def:untwisted-finite-gauge-partition}
For a closed \(D\)-manifold \(M\), the untwisted finite \(G\)-gauge theory
partition function is
\begin{equation}
  Z_G(M)
  =
  |\operatorname{Bun}_G(M)|
  =
  \sum_{[P]\in\pi_0\operatorname{Bun}_G(M)}
  \frac{1}{|\operatorname{Aut}(P)|}.
  \label{eq:untwisted-finite-gauge-partition}
\end{equation}
\end{definition}

If \(M\) is connected and \(m\in M\) is a basepoint, a principal
\(G\)-bundle is classified by a monodromy homomorphism
\[
  \rho:\pi_1(M,m)\to G
\]
up to conjugation by \(G\).  The automorphism group of the corresponding
bundle is
\[
  C_G(\rho)
  =
  \{h\in G: h\rho(\gamma)h^{-1}=\rho(\gamma)
  \ \hbox{for all }\gamma\in\pi_1(M,m)\}.
\]

For connected closed \(M\),
\begin{equation}
  Z_G(M)
  =
  \frac{|\operatorname{Hom}(\pi_1(M),G)|}{|G|}.
  \label{eq:finite-gauge-connected-formula}
\end{equation}
The groupoid \(\operatorname{Bun}_G(M)\) is equivalent to the action groupoid
\[
  \operatorname{Hom}(\pi_1(M),G)//G,
\]
where \(G\) acts by conjugation.  The stabilizer of \(\rho\) is
\(C_G(\rho)\), which is the automorphism group of the corresponding bundle.
The action-groupoid cardinality calculation gives
\eqref{eq:finite-gauge-connected-formula}.
This argument uses connectedness to choose one basepoint and identify every
flat finite \(G\)-bundle with a homomorphism \(\pi_1(M,m)\to G\) modulo
change of trivialization at the basepoint, namely conjugation by \(G\).

For a disconnected \(M=\bigsqcup_{\alpha=1}^c M_\alpha\), groupoid
cardinality is multiplicative under disjoint union:
\[
  Z_G(M)
  =
  \prod_{\alpha=1}^c Z_G(M_\alpha).
\]
The empty manifold has value \(1\), as required for a symmetric monoidal
field theory.

\section{State Spaces and Push-Pull Along Spans}

Let \(\Sigma\) be a closed \((D-1)\)-manifold.  The boundary field groupoid in
Definition~\ref{def:finite-gauge-state-sum-datum} is
\[
  \mathcal B_\Sigma=\operatorname{Bun}_G(\Sigma).
\]
In the untwisted theory the state space is
\begin{equation}
  \mathcal H_G(\Sigma)
  =
  \operatorname{Fun}(\pi_0\mathcal B_\Sigma,\C),
  \label{eq:finite-gauge-state-space}
\end{equation}
the vector space of complex-valued functions on isomorphism classes of
boundary bundles.  Its canonical bilinear pairing is the groupoid integral
\begin{equation}
  \langle f_1,f_2\rangle_\Sigma
  =
  \int_{\mathcal B_\Sigma} f_1(Q)f_2(Q)
  =
  \sum_{[Q]}
  \frac{f_1(Q)f_2(Q)}{|\operatorname{Aut}(Q)|}.
  \label{eq:finite-gauge-boundary-pairing}
\end{equation}
The pairing identifies \(\mathcal H_G(\bar\Sigma)\) with the dual of
\(\mathcal H_G(\Sigma)\), in agreement with
Proposition~\ref{prop:tqft-state-space-duality}.

If
\[
  M:\Sigma_0\to\Sigma_1
\]
is a collared bordism, restriction to the boundary gives a span of finite
groupoids
\begin{equation}
  \mathcal B_{\Sigma_0}
  \xleftarrow{\ r_0\ }
  \operatorname{Bun}_G(M)
  \xrightarrow{\ r_1\ }
  \mathcal B_{\Sigma_1}.
  \label{eq:finite-gauge-bordism-span}
\end{equation}
The bordism operator is the push-pull map
\begin{equation}
  Z_G(M)
  =
  (r_1)_!\,r_0^*
  :
  \mathcal H_G(\Sigma_0)\to\mathcal H_G(\Sigma_1).
  \label{eq:finite-gauge-push-pull}
\end{equation}
Here \(r_0^*\) is pullback of functions, and \((r_1)_!\) is integration over
the finite homotopy fiber of \(r_1\).

Concretely, if \(Q_1\in\mathcal B_{\Sigma_1}\), then
\begin{equation}
  \bigl(Z_G(M)f\bigr)(Q_1)
  =
  \int_{P\in\operatorname{Bun}_G(M)_{Q_1}}
  f(r_0P),
  \label{eq:finite-gauge-pushforward-concrete}
\end{equation}
where \(\operatorname{Bun}_G(M)_{Q_1}\) is the groupoid of pairs
\((P,\alpha)\) with \(P\in\operatorname{Bun}_G(M)\) and an isomorphism
\(\alpha:r_1P\simeq Q_1\).

\begin{lemma}[Finite groupoid Fubini with automorphism weights]
\label{lem:finite-groupoid-fubini}
Let
\[
  p:E\to B
\]
be a functor between finite groupoids, and let \(F:E\to\C\) be constant on
isomorphism classes.  For \(b\in B\), let \(E_b\) be the homotopy fiber
groupoid whose objects are pairs \((e,\phi)\), with \(e\in E\) and
\(\phi:p(e)\simeq b\) in \(B\).  Then
\begin{equation}
  \int_E F
  =
  \int_{b\in B}\int_{E_b} F.
  \label{eq:finite-groupoid-fubini}
\end{equation}
Consequently, if \(X\xrightarrow{f}Y\xrightarrow{g}Z\) are finite groupoids,
then
\begin{equation}
  (g\circ f)_!F=g_!f_!F.
  \label{eq:finite-groupoid-pushforward-composition}
\end{equation}
\end{lemma}

\begin{proof}
Choose representatives \(b\) for the isomorphism classes of \(B\).  The
right hand side of \eqref{eq:finite-groupoid-fubini} is
\[
  \sum_{[b]\in\pi_0B}
  \frac{1}{|\operatorname{Aut}_B(b)|}
  \sum_{[(e,\phi)]\in\pi_0E_b}
  \frac{F(e)}{|\operatorname{Aut}_{E_b}(e,\phi)|}.
\]
We compare the contribution of one isomorphism class \([e]\in\pi_0E\).
After replacing \(b\) by an isomorphic representative, assume
\[
  b=p(e).
\]
Let
\[
  I_e=\operatorname{im}\bigl(\operatorname{Aut}_E(e)\to
  \operatorname{Aut}_B(p(e))\bigr),
  \qquad
  K_e=\ker\bigl(\operatorname{Aut}_E(e)\to
  \operatorname{Aut}_B(p(e))\bigr).
\]
The possible identifications \(\phi:p(e)\simeq b\) form a torsor for
\(\operatorname{Aut}_B(b)\).  Passing to the homotopy fiber divides this
torsor by the image \(I_e\).  Hence the class \([e]\) gives
\[
  \frac{|\operatorname{Aut}_B(b)|}{|I_e|}
\]
isomorphism classes in \(E_b\), and the automorphism group of each such
fiber object is \(K_e\).  Its total contribution to the iterated integral is
\[
  \frac{1}{|\operatorname{Aut}_B(b)|}
  \frac{|\operatorname{Aut}_B(b)|}{|I_e|}
  \frac{F(e)}{|K_e|}
  =
  \frac{F(e)}{|I_e|\,|K_e|}
  =
  \frac{F(e)}{|\operatorname{Aut}_E(e)|}.
\]
This is exactly the contribution of \([e]\) to \(\int_EF\).  Summing over
\(\pi_0E\) proves \eqref{eq:finite-groupoid-fubini}.

For pushforwards, the homotopy fiber of \(g\circ f\) over \(z\in Z\) is
equivalent to the homotopy fiber product
\[
  X\times_Y Y_z.
\]
Applying \eqref{eq:finite-groupoid-fubini} to the projection
\[
  X\times_Y Y_z\to Y_z
\]
gives the equality between integration over this single homotopy fiber and
the iterated integration first over \(X_y\) and then over \(Y_z\).  This is
\eqref{eq:finite-groupoid-pushforward-composition}.
\end{proof}

\begin{theorem}[Finite groupoid gluing]
\label{thm:finite-groupoid-gluing}
Let \(M_{01}:\Sigma_0\to\Sigma_1\) and
\(M_{12}:\Sigma_1\to\Sigma_2\) be collared bordisms.  The natural
restriction functor gives an equivalence of finite groupoids
\begin{equation}
  \operatorname{Bun}_G(M_{12}\circ M_{01})
  \simeq
  \operatorname{Bun}_G(M_{01})
  \times_{\operatorname{Bun}_G(\Sigma_1)}
  \operatorname{Bun}_G(M_{12}),
  \label{eq:bun-gluing-homotopy-pullback}
\end{equation}
where the fiber product is the homotopy fiber product of groupoids.  With
the push-pull definition \eqref{eq:finite-gauge-push-pull},
\begin{equation}
  Z_G(M_{12}\circ M_{01})
  =
  Z_G(M_{12})\,Z_G(M_{01}).
  \label{eq:finite-gauge-gluing}
\end{equation}
\end{theorem}

\begin{proof}
The equivalence \eqref{eq:bun-gluing-homotopy-pullback} is descent for
principal bundles over the collar cover
\[
  M_{12}\circ M_{01}=M_{01}\cup_{\Sigma_1}M_{12}.
\]
An object on the right is a bundle \(P_{01}\to M_{01}\), a bundle
\(P_{12}\to M_{12}\), and an isomorphism
\[
  P_{01}|_{\Sigma_1}\simeq P_{12}|_{\Sigma_1}.
\]
Gluing the total spaces along this isomorphism gives a principal bundle on
the glued manifold.  Conversely, restricting a bundle on the glued manifold
gives such data.  A bundle isomorphism on the glued manifold restricts to a
pair of isomorphisms on \(M_{01}\) and \(M_{12}\) compatible with the chosen
boundary identification; conversely, any compatible pair glues uniquely to an
isomorphism of the glued bundles.  Thus the equivalence includes the
morphism groupoids, not only the object sets.

The bordism operator is the linear map associated to the restriction span.
Composition of such span operators is composition of finite-groupoid
pullbacks and pushforwards.  By Lemma~\ref{lem:finite-groupoid-fubini}, the
pushforward along a composite finite-groupoid map is the iterated
pushforward, with the automorphism denominators matching by the
image-kernel calculation in the lemma.  Combining this Fubini identity with
the descent equivalence \eqref{eq:bun-gluing-homotopy-pullback} gives
\[
  Z_G(M_{12}\circ M_{01})
  =
  Z_G(M_{12})\,Z_G(M_{01}),
\]
as an equality of linear maps
\(\mathcal H_G(\Sigma_0)\to\mathcal H_G(\Sigma_2)\).
\end{proof}

Figure~\ref{fig:finite-gauge-gluing-span} records the span-level gluing proof:
restriction from bulk \(G\)-bundles to boundary \(G\)-bundles composes by
homotopy pullback, and groupoid cardinality supplies the Fubini identity.

\begin{figure}[H]
\centering
\begin{tikzpicture}[>=Stealth,scale=0.95]
  \node (B0) at (0,0) {\(\operatorname{Bun}_G(\Sigma_0)\)};
  \node (X) at (3.2,1.15) {\(\operatorname{Bun}_G(M_{01})\)};
  \node (B1) at (6.4,0) {\(\operatorname{Bun}_G(\Sigma_1)\)};
  \node (Y) at (9.6,1.15) {\(\operatorname{Bun}_G(M_{12})\)};
  \node (B2) at (12.8,0) {\(\operatorname{Bun}_G(\Sigma_2)\)};
  \draw[<-] (B0) -- (X);
  \draw[->] (X) -- (B1);
  \draw[<-] (B1) -- (Y);
  \draw[->] (Y) -- (B2);
  \node at (6.4,2.25)
    {\(\operatorname{Bun}_G(M_{01})\times_{\operatorname{Bun}_G(\Sigma_1)}
      \operatorname{Bun}_G(M_{12})\)};
  \draw[dashed,->] (6.4,2.0) -- (4.2,1.35);
  \draw[dashed,->] (6.4,2.0) -- (8.6,1.35);
\end{tikzpicture}
\caption{Gluing finite gauge theory is composition of spans of finite
groupoids.  The middle groupoid is retained as a homotopy fiber product, so
automorphisms of the intermediate boundary bundle are divided out exactly
once.}
\label{fig:finite-gauge-gluing-span}
\end{figure}

\section{Triangulated State Sums}

Let \(K\) be a finite triangulation of a closed oriented \(D\)-manifold
\(M\), equipped with a branching, meaning an ordering of vertices on every
simplex compatible on faces.  A lattice \(G\)-connection assigns an element
\[
  g_{ij}\in G
\]
to every oriented edge \(i\to j\), with \(g_{ji}=g_{ij}^{-1}\).  It is flat
when
\begin{equation}
  g_{ij}g_{jk}=g_{ik}
  \label{eq:finite-gauge-flatness}
\end{equation}
on every oriented triangle.  Gauge transformations are vertex labels
\[
  h_i\in G,
\]
acting by
\begin{equation}
  g_{ij}\longmapsto h_i g_{ij}h_j^{-1}.
  \label{eq:finite-gauge-lattice-gauge-transformation}
\end{equation}
Let \(\operatorname{Flat}_G(K)\) be the finite set of flat edge labelings.
The action groupoid
\[
  \operatorname{Flat}_G(K)//G^{K_0}
\]
is equivalent to \(\operatorname{Bun}_G(M)\).  Therefore
\begin{equation}
  Z_G(M)
  =
  \frac{|\operatorname{Flat}_G(K)|}{|G|^{|K_0|}}.
  \label{eq:finite-gauge-lattice-state-sum}
\end{equation}

For connected \(K\), choose a spanning tree \(T\) in the one-skeleton.  Every
flat connection is gauge equivalent to a unique flat connection with
\[
  g_{ij}=1
  \qquad\hbox{for every edge }(ij)\in T,
\]
up to the remaining constant conjugation by \(G\).  This gauge fixing gives
the equality between \eqref{eq:finite-gauge-lattice-state-sum} and
\eqref{eq:finite-gauge-connected-formula}.

\section{Dijkgraaf--Witten Twists}

The Dijkgraaf--Witten entry \([\omega]\) in
Definition~\ref{def:finite-gauge-state-sum-datum} is represented by
\[
  \omega\in Z^D(BG,U(1))
\]
normalized as a group cocycle.  In an inhomogeneous cochain convention this
means a function
\[
  \omega:G^D\to U(1)
\]
which is \(1\) when any argument is the identity and satisfies
\begin{equation}
  (\delta\omega)(g_1,\ldots,g_{D+1})=1.
  \label{eq:dw-cocycle-condition}
\end{equation}
For example, for \(D=2\),
\begin{equation}
  \omega(g_2,g_3)\,
  \omega(g_1g_2,g_3)^{-1}\,
  \omega(g_1,g_2g_3)\,
  \omega(g_1,g_2)^{-1}
  =
  1.
  \label{eq:dw-2-cocycle-condition}
\end{equation}

For a branched oriented \(D\)-simplex
\(\sigma=[i_0\cdots i_D]\), define
\begin{equation}
  \omega_\sigma(g)
  =
  \omega(g_{i_0i_1},g_{i_1i_2},\ldots,g_{i_{D-1}i_D})^{\varepsilon(\sigma)},
  \label{eq:dw-simplex-weight}
\end{equation}
where \(\varepsilon(\sigma)=+1\) if the branching order agrees with the
orientation of \(M\) on \(\sigma\), and \(-1\) otherwise.  The twisted
state-sum is
\begin{equation}
  Z_{G,\omega}(M;K)
  =
  \frac{1}{|G|^{|K_0|}}
  \sum_{g\in\operatorname{Flat}_G(K)}
  \prod_{\sigma^D\in K_D}\omega_\sigma(g).
  \label{eq:dw-triangulated-state-sum}
\end{equation}

\begin{proposition}[Triangulation independence of the cocycle weight]
\label{prop:dw-triangulation-independence}
For closed \(M\), the quantity
\eqref{eq:dw-triangulated-state-sum} is invariant under branching-preserving
Pachner moves and changes by the expected boundary coboundary factor if
\(\omega\) is replaced by \(\omega\,\delta\lambda\).  Hence on closed
manifolds it depends only on the cohomology class
\[
  [\omega]\in H^D(BG,U(1)).
\]
\end{proposition}

\begin{proof}
A flat edge labeling of a branched triangulation is equivalently a simplicial
map from the triangulation to the nerve model of \(BG\): the ordered edge
labels on a simplex are composable group elements, and flatness says that
the label on a longer edge is their product.  The factor
\(\omega_\sigma(g)\) is therefore the pullback of the normalized
\(D\)-cochain \(\omega\) evaluated on the oriented \(D\)-simplex \(\sigma\).

A branching-preserving Pachner move replaces a collection of oriented
\(D\)-faces in the boundary of one branched \((D+1)\)-simplex by the
complementary collection of \(D\)-faces.  For a fixed flat labeling, all
boundary labels are restrictions of one \((D+1)\)-simplex label
\((g_1,\ldots,g_{D+1})\).  The ratio between the two products of
\(\omega\)-weights is the alternating product over the \(D\)-faces of this
\((D+1)\)-simplex, namely
\[
  (\delta\omega)(g_1,\ldots,g_{D+1}),
\]
where the exponent sign records whether the face orientation agrees with the
boundary orientation.  The cocycle condition
\eqref{eq:dw-cocycle-condition} makes this ratio equal to \(1\).  Since a
Pachner move is local and gives a bijection between flat labelings that
agree outside the moved ball, the whole state-sum is unchanged.

Now replace \(\omega\) by \(\omega\,\delta\lambda\), with \(\lambda\) a
\((D-1)\)-cochain on \(G\).  On each \(D\)-simplex the extra factor is the
alternating product of the \(\lambda\)-weights on its \((D-1)\)-faces.  When
one multiplies over all \(D\)-simplices, every interior \((D-1)\)-face
appears twice with opposite induced orientations and cancels.  The surviving
product is supported on \(\partial M\).  Thus a closed \(M\) has no residual
factor, while a manifold with boundary changes by exactly the boundary
coboundary term.  The closed state-sum therefore depends only on
\([\omega]\in H^D(BG,U(1))\).
\end{proof}

The intrinsic form of the same definition is the following.  A bundle
\(P\to M\) has a classifying map
\[
  f_P:M\to BG
\]
well-defined up to homotopy.  The cocycle supplies a phase
\[
  \exp\!\left(2\pi\ii\int_M f_P^*\omega\right).
\]
Then
\begin{equation}
  Z_{G,\omega}(M)
  =
  \sum_{[P]\in\pi_0\operatorname{Bun}_G(M)}
  \frac{1}{|\operatorname{Aut}(P)|}
  \exp\!\left(2\pi\ii\int_M f_P^*\omega\right).
  \label{eq:dw-intrinsic-partition}
\end{equation}
Equation~\eqref{eq:dw-triangulated-state-sum} is the simplicial
representative of \eqref{eq:dw-intrinsic-partition}.

On a manifold with boundary, the same transgression produces a line over
\(\operatorname{Bun}_G(\partial M)\).  The state space is therefore not
ordinary functions but sections of that line:
\begin{equation}
  \mathcal H_{G,\omega}(\Sigma)
  =
  \Gamma\bigl(\operatorname{Bun}_G(\Sigma),\mathcal L_\omega(\Sigma)\bigr).
  \label{eq:dw-twisted-state-space}
\end{equation}
The gluing theorem continues to hold because the two transgressed boundary
lines on a cut hypersurface are dual and contract under gluing.

\section{Explicit Closed-Surface Partition Functions}

Let \(\Sigma_g\) be a closed connected oriented surface of genus \(g\).  Its
fundamental group has presentation
\[
  \pi_1(\Sigma_g)
  =
  \left\langle
    a_1,b_1,\ldots,a_g,b_g
    \ \middle|\
    \prod_{r=1}^g[a_r,b_r]=1
  \right\rangle .
\]
Thus
\begin{equation}
  Z_G(\Sigma_g)
  =
  \frac{1}{|G|}
  \#\left\{
    (A_r,B_r)_{r=1}^g\in G^{2g}:
    \prod_{r=1}^g[A_r,B_r]=1
  \right\}.
  \label{eq:finite-gauge-surface-hom-count}
\end{equation}

\begin{proposition}[Character formula for the \(2D\) untwisted theory]
\label{prop:finite-gauge-surface-character-formula}
Let \(\operatorname{Irr}(G)\) be the set of irreducible complex
representations of \(G\), and let \(d_R=\dim R\).  Then
\begin{equation}
  Z_G(\Sigma_g)
  =
  |G|^{2g-2}
  \sum_{R\in\operatorname{Irr}(G)}
  d_R^{\,2-2g}
  =
  \sum_{R\in\operatorname{Irr}(G)}
  \left(\frac{d_R}{|G|}\right)^{2-2g}.
  \label{eq:finite-gauge-surface-character-formula}
\end{equation}
\end{proposition}

\begin{proof}
Let
\[
  \delta_e(x)=
  \begin{cases}
  1,&x=e,\\
  0,&x\ne e
  \end{cases}
\]
be the delta function at the identity.  The character expansion of the
central delta function is
\begin{equation}
  \delta_e(x)
  =
  \frac{1}{|G|}
  \sum_{R\in\operatorname{Irr}(G)}d_R\,\chi_R(x),
  \label{eq:finite-group-delta-character-expansion}
\end{equation}
which follows from the orthogonality of irreducible characters.  Define the
commutator-sum operator on class functions by
\[
  (Tf)(x)
  =
  \sum_{a,b\in G} f(x[a,b]).
\]
Since \(T\) commutes with conjugation and convolution by central functions,
Schur orthogonality gives
\begin{equation}
  T\chi_R
  =
  \frac{|G|^2}{d_R^2}\,\chi_R.
  \label{eq:commutator-sum-character-eigenvalue}
\end{equation}
To see this, write the representation matrix also as \(R(\cdot)\).  Schur
averaging says that for every endomorphism \(A\) of \(R\),
\begin{equation}
  \sum_{b\in G} R(b)A R(b)^{-1}
  =
  \frac{|G|\,\operatorname{tr}_R A}{d_R}\,{\bf 1}_R.
  \label{eq:finite-schur-averaging}
\end{equation}
With \(A=R(a)^{-1}\),
\[
  \sum_{b\in G} R([a,b])
  =
  R(a)\sum_{b\in G}R(b)R(a)^{-1}R(b)^{-1}
  =
  \frac{|G|\chi_R(a^{-1})}{d_R}R(a).
\]
The central-idempotent form of character orthogonality gives
\[
  \sum_{a\in G}\chi_R(a^{-1})R(a)
  =
  \frac{|G|}{d_R}\,{\bf 1}_R.
\]
Therefore
\[
  \sum_{a,b\in G}R([a,b])
  =
  \frac{|G|^2}{d_R^2}\,{\bf 1}_R,
\]
and multiplying by \(R(x)\) and taking the trace proves
\eqref{eq:commutator-sum-character-eigenvalue}.

The number of solutions of
\(\prod_{r=1}^g[A_r,B_r]=1\) is \(T^g\delta_e(e)\).  Applying
\eqref{eq:finite-group-delta-character-expansion} and
\eqref{eq:commutator-sum-character-eigenvalue} gives
\[
  \#\operatorname{Hom}(\pi_1(\Sigma_g),G)
  =
  |G|^{2g-1}
  \sum_R d_R^{\,2-2g}.
\]
Dividing by the residual global gauge volume \(|G|\) gives
\eqref{eq:finite-gauge-surface-character-formula}.
\end{proof}

For \(g=0\), the formula gives
\[
  Z_G(S^2)=\frac{1}{|G|}
\]
because \(\sum_Rd_R^2=|G|\).  For \(g=1\),
\[
  Z_G(T^2)=|\operatorname{Irr}(G)|,
\]
which is also the number of conjugacy classes of \(G\).  If \(G=A\) is
finite abelian, all irreducible representations are one-dimensional, so
\begin{equation}
  Z_A(\Sigma_g)=|A|^{2g-1}.
  \label{eq:finite-abelian-surface-partition}
\end{equation}

\section{Circle State Space and the Two-Dimensional Algebra}

For the circle,
\[
  \operatorname{Bun}_G(S^1)\simeq G//G,
\]
where \(G\) acts on itself by conjugation.  Therefore
\begin{equation}
  \mathcal H_G(S^1)
  =
  \{\hbox{class functions }G\to\C\}.
  \label{eq:finite-gauge-circle-class-functions}
\end{equation}
The pairing \eqref{eq:finite-gauge-boundary-pairing} becomes
\begin{equation}
  \langle f_1,f_2\rangle_{S^1}
  =
  \frac{1}{|G|}\sum_{g\in G} f_1(g)f_2(g).
  \label{eq:finite-gauge-circle-pairing}
\end{equation}
The pair of pants defines a commutative Frobenius algebra on class
functions.  In the orientation convention where the outgoing holonomy is the
product of the incoming holonomies, the product is convolution
\begin{equation}
  (f_1*f_2)(k)
  =
  \sum_{\substack{a,b\in G\\ab=k}}
  f_1(a)f_2(b).
  \label{eq:finite-gauge-class-function-convolution}
\end{equation}
Associativity is associativity of group multiplication plus finite Fubini.
Commutativity on class functions follows by the change of variables
\[
  (a,b)\mapsto (aba^{-1},a),
\]
which preserves the product \(ab\) up to conjugacy; class functions cannot
distinguish conjugate outgoing holonomies.  The counit is the disk amplitude,
and the nondegenerate pairing is \eqref{eq:finite-gauge-circle-pairing}.
This identifies the ordinary \(2D\) finite gauge theory with the Frobenius
algebra construction of
Theorem~\ref{thm:2d-oriented-tqft-frobenius}.

\section{Three-Dimensional Finite Gauge Theory}

For a closed connected \(3\)-manifold \(M\), the untwisted partition function
is again
\[
  Z_G(M)=\frac{|\operatorname{Hom}(\pi_1(M),G)|}{|G|}.
\]
In particular,
\begin{equation}
  Z_G(T^3)
  =
  \frac{1}{|G|}
  \#\{(a,b,c)\in G^3:
  ab=ba,\ ac=ca,\ bc=cb\}.
  \label{eq:finite-gauge-t3-partition}
\end{equation}
The state space on \(T^2\) is the vector space of functions on commuting
pairs \((a,b)\) modulo simultaneous conjugation.  The extended
three-dimensional theory assigns to the circle a category.  In the untwisted
case this category is equivalent to representations of the Drinfeld double
\[
  D(G),
\]
and in the \(\omega\)-twisted case to representations of the twisted quantum
double.  Equivalently, the line operators form the Drinfeld center
\[
  Z(\operatorname{Vect}^{\omega}_G).
\]
The simple objects are labeled by a conjugacy class \(\mathcal C\subset G\)
and an irreducible projective representation of the centralizer \(C_G(c)\),
with the projective \(2\)-cocycle obtained by transgressing \(\omega\) to the
flux \(c\in\mathcal C\).  Fusion and braiding are finite tensor-category
operations.  This is a theorem of finite tensor categories, but the
path-integral origin is already visible: a line insertion fixes magnetic
monodromy around a small linking circle, while electric labels are
representations of the residual stabilizer of that monodromy.

\section{Defects, Boundaries, and Anomaly Inflow}

Finite gauge theories also provide exact models of defects.  A
codimension-one defect between \(G\)- and \(H\)-gauge theories is described,
in the finite path-integral language, by a finite correspondence of boundary
field groupoids
\begin{equation}
  \operatorname{Bun}_G(\Sigma)
  \xleftarrow{\ }
  \mathcal D(\Sigma)
  \xrightarrow{\ }
  \operatorname{Bun}_H(\Sigma),
  \label{eq:finite-gauge-defect-correspondence}
\end{equation}
possibly decorated by a cocycle line.  A concrete subgroup defect is obtained
from
\[
  K\subset G\times H,
\]
by requiring the two gauge fields to reduce on the wall to a common
principal \(K\)-bundle.  Composition of such defects is again a homotopy
fiber product of finite groupoids.

In a \(D\)-dimensional theory with finite global symmetry \(G\), an anomaly
is represented by a class
\[
  [\alpha]\in H^{D+1}(BG,U(1)).
\]
For a background \(G\)-bundle \(A\) on a closed \(D\)-manifold \(M\), the
would-be anomalous partition function is not a number canonically attached
to \(A\).  It is a vector in the transgressed line determined by \(\alpha\).
Keeping a \((D+1)\)-dimensional bulk with action \(\alpha\) makes the
combined bulk-boundary amplitude gauge invariant.  Gauging the boundary
symmetry alone requires a trivialization of this anomaly line.  This is the
finite model of the relative field theory language in
Definition~\ref{def:anomalous-relative-tqft}.

\begin{openproblem}[Extended finite models]
\label{op:extended-finite-models}
Construct extended finite-gauge TQFTs with boundaries, defects, junctions,
twists, and anomaly inflow as functors on a chosen bordism
\((\infty,D)\)-category, including a direct comparison between the
higher-categorical functor and the finite groupoid state-sum formulas in
this chapter.
\end{openproblem}
