\chapter{Conformal Killing Vectors and the Conformal Group}
\label{ch:volume-iii-conformal-killing-vectors}
\ConventionsEuclideanCFT


\section{The Conformal Killing Equation}

Work first in flat Euclidean space \((\R^D,\delta_{\mu\nu})\) with
\(D\ge3\).  Cartesian coordinates are denoted by \(x^\mu\), and repeated
indices are contracted with \(\delta_{\mu\nu}\).  A smooth vector field
\(\epsilon^\mu(x)\) generates an
infinitesimal coordinate transformation
\[
  x^\mu\longmapsto x^\mu+\epsilon^\mu(x).
\]
It is conformal when the change of the metric is proportional to the metric:
\[
  \mathcal L_\epsilon\delta_{\mu\nu}
  =
  \partial_\mu\epsilon_\nu+\partial_\nu\epsilon_\mu
  =
  2\sigma_\epsilon(x)\delta_{\mu\nu}
\]
for some scalar function \(\sigma_\epsilon\).  Taking the trace gives
\[
  \sigma_\epsilon(x)
  =
  \frac1D\partial_\rho\epsilon^\rho(x).
\]
Thus the conformal Killing equation is
\begin{equation}
  \partial_\mu\epsilon_\nu+\partial_\nu\epsilon_\mu
  -
  \frac2D\delta_{\mu\nu}\partial_\rho\epsilon^\rho
  =
  0.
  \label{eq:conformal-killing-equation}
\end{equation}

If \(T_{\mu\nu}\) is symmetric, conserved, and traceless, then
\[
  J^\mu_\epsilon(x)=T^{\mu}{}_{\nu}(x)\epsilon^\nu(x)
\]
is conserved:
\[
  \partial_\mu J^\mu_\epsilon
  =
  (\partial_\mu T^{\mu}{}_{\nu})\epsilon^\nu
  +
  T^{\mu\nu}\partial_\mu\epsilon_\nu
  =
  \frac12T^{\mu\nu}
  (\partial_\mu\epsilon_\nu+\partial_\nu\epsilon_\mu)
  =
  \sigma_\epsilon T^\mu{}_\mu
  =
  0.
\]
The corresponding charge on a closed codimension-one surface \(\Sigma\) is
\[
  Q_\epsilon(\Sigma)
  =
  \int_\Sigma \dd\sigma\,n_\mu J^\mu_\epsilon,
\]
where \(n_\mu\) is the unit normal.  Conservation makes the charge invariant
under deformations of \(\Sigma\) that do not cross operator insertions.
After analytic continuation to Lorentzian signature this formula is read on a
spacelike Cauchy surface \(\Sigma\).  The measure \(\dd\sigma\) is the induced
spatial volume form, \(n^\mu\) is the future-directed unit timelike normal, and
\begin{equation}
  Q_\epsilon(\Sigma)
  =
  \int_\Sigma \dd\sigma\,n_\mu T^\mu{}_\nu\epsilon^\nu .
  \label{eq:lorentzian-conformal-charge}
\end{equation}
If \(\Sigma_1\) and \(\Sigma_2\) bound a spacetime region \(V\) containing no
operator insertions, Stokes' theorem gives
\[
  Q_\epsilon(\Sigma_2)-Q_\epsilon(\Sigma_1)
  =
  \int_V \dd^D x\,\partial_\mu J^\mu_\epsilon
  =
  0.
\]
This is the precise sense in which the charge depends on the homology class of
the surface relative to insertions.

\begin{figure}[H]
  \centering
  \resizebox{0.72\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \draw[-{Latex[length=2mm]},thick] (-4.0,-1.25)--(-4.0,2.4)
      node[above] {time};
    \draw[-{Latex[length=2mm]},thick] (-4.25,-1.0)--(4.1,-1.0)
      node[right] {space};
    \draw[blue!70!black,very thick]
      (-3.4,0.0) .. controls (-2.0,0.45) and (-0.7,-0.2) ..
      (0.65,0.25) .. controls (1.7,0.60) and (2.55,0.15) .. (3.4,0.50);
    \node[blue!70!black] at (3.75,0.60) {\(\Sigma\)};
    \draw[fill=purple!65!black,draw=purple!65!black]
      (-0.45,0.22) circle (0.055);
    \draw[-{Latex[length=2mm]},very thick,purple!65!black]
      (-0.45,0.22)--(-0.45,1.35);
    \node[purple!65!black,align=center] at (0.25,1.45)
      {\(n^\mu\)\\timelike normal};
    \node[align=center] at (0,-1.75)
      {\(Q_\epsilon(\Sigma)=\displaystyle\int_\Sigma
      \dd\sigma\,n_\mu T^\mu{}_\nu\epsilon^\nu\)};
  \end{tikzpicture}
  }
  \caption{The Lorentzian conformal charge is the flux of the stress-tensor
  current through a spacelike slice.  Current conservation and Stokes' theorem
  give independence under deformations of the slice through empty spacetime.}
  \label{fig:conformal-charge-spacelike-slice}
\end{figure}

\section{Solutions in Dimension at Least Three}

For \(D\ge3\), differentiating
\eqref{eq:conformal-killing-equation} and combining permutations of the
indices gives
\[
  \partial_\mu\partial_\nu\epsilon_\rho
  =
  \delta_{\rho\nu}\partial_\mu\sigma_\epsilon
  +
  \delta_{\rho\mu}\partial_\nu\sigma_\epsilon
  -
  \delta_{\mu\nu}\partial_\rho\sigma_\epsilon .
\]
Taking another derivative and antisymmetrizing shows
\[
  \partial_\mu\partial_\nu\partial_\rho\sigma_\epsilon=0.
\]
Hence \(\sigma_\epsilon\) is at most linear in \(x\), and
\(\epsilon^\mu\) is at most quadratic.  The general solution is
\begin{equation}
  \epsilon^\mu(x)
  =
  a^\mu
  +
  \omega^\mu{}_\nu x^\nu
  +
  \lambda x^\mu
  +
  c^\mu x^2-2(c\cdot x)x^\mu,
  \label{eq:ckv-general-solution}
\end{equation}
where \(a^\mu\), \(c^\mu\), and \(\lambda\) are constants and
\(\omega_{\mu\nu}=-\omega_{\nu\mu}\).  These terms generate translations,
rotations, dilatations, and special conformal transformations, respectively.
Equivalently, the Lorentz and dilatation terms may be combined as
\(b^\mu{}_\nu x^\nu\), where \(b_{\mu\nu}\) is either antisymmetric or
proportional to \(\delta_{\mu\nu}\).

The associated differential operators are
\begin{equation}
  P_\mu=\partial_\mu,\qquad
  M_{\mu\nu}=x_\mu\partial_\nu-x_\nu\partial_\mu,
  \qquad
  D=x^\mu\partial_\mu,
  \label{eq:conformal-differential-generators-1}
\end{equation}
and
\begin{equation}
  K_\mu=x^2\partial_\mu-2x_\mu x^\rho\partial_\rho .
  \label{eq:conformal-differential-generators-2}
\end{equation}
The signs in \eqref{eq:conformal-differential-generators-2} are chosen so
that the finite special conformal transformation below has parameter \(c^\mu\)
with denominator \(1+2c\cdot x+c^2x^2\).

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,2.25) {conformal Killing vector};
    \draw[rounded corners=4pt,thick] (-2.0,0.45) rectangle (2.0,1.55);
    \node at (0,1.00) {\(\epsilon^\mu(x)\)};

    \draw[-{Latex[length=2mm]},thick] (2.25,1.0)--(3.45,1.0);

    \begin{scope}[shift={(5.35,0)}]
      \node[font=\small] at (0,2.25) {finite-dimensional data};
      \draw[rounded corners=4pt,thick] (-2.55,0.35) rectangle (2.55,1.65);
      \node at (0,1.27) {\(a^\mu,\ \omega_{\mu\nu},\ \lambda,\ c^\mu\)};
      \node[align=center] at (0,0.72)
        {translation, rotation,\\dilatation, special conformal};
    \end{scope}

    \draw[-{Latex[length=2mm]},thick] (7.95,1.0)--(9.15,1.0);

    \begin{scope}[shift={(11.0,0)}]
      \node[font=\small] at (0,2.25) {charges};
      \draw[rounded corners=4pt,thick] (-1.75,0.45) rectangle (1.75,1.55);
      \node at (0,1.00) {\(Q_\epsilon=\int_\Sigma n_\mu T^\mu{}_\nu\epsilon^\nu\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{In \(D\ge3\), a conformal Killing vector is determined by finitely
  many constants.  The stress tensor converts such a vector field into a
  conserved charge.}
  \label{fig:ckv-data-to-charge}
\end{figure}

\section{Charges and Vector Fields}

Let \(U_\epsilon(s)\) denote the one-parameter unitary family generated by
\(Q_\epsilon\):
\[
  U_\epsilon(s)=\exp(\ii s Q_\epsilon).
\]
The normalization of \(Q_\epsilon\) is fixed by translations and Lorentz
transformations, whose charges are already part of the Poincare
representation of the QFT.  With this normalization, infinitesimal coordinate
transformations compose through the ordinary Lie bracket of vector fields,
\begin{equation}
  [\epsilon_1,\epsilon_2]^\mu
  =
  \epsilon_1^\nu\partial_\nu\epsilon_2^\mu
  -
  \epsilon_2^\nu\partial_\nu\epsilon_1^\mu .
  \label{eq:ckv-lie-bracket}
\end{equation}
For the charges the corresponding commutator is
\begin{equation}
  [Q_{\epsilon_1},Q_{\epsilon_2}]
  =
  -\ii Q_{[\epsilon_1,\epsilon_2]} .
  \label{eq:charge-vector-field-commutator}
\end{equation}
Equation \eqref{eq:charge-vector-field-commutator} is an operator statement on
a common invariant domain, or equivalently a Ward identity for correlation
functions with separated insertions.  Possible central extensions or
Schwinger terms are local contact terms; in dimensions \(D\ge3\) the global
finite-dimensional conformal algebra has no central extension.

\section{Finite Transformations}

A finite conformal transformation is a diffeomorphism
\[
  f:\R^D\supset U\longrightarrow f(U)\subset\R^D
\]
whose Jacobian satisfies
\begin{equation}
  \frac{\partial f^\rho}{\partial x^\mu}
  \frac{\partial f^\sigma}{\partial x^\nu}
  \delta_{\rho\sigma}
  =
  \Omega_f(x)^2\delta_{\mu\nu},
  \qquad
  \Omega_f(x)>0.
  \label{eq:finite-conformal-map}
\end{equation}
For a transformation close to the identity,
\[
  f^\mu(x)=x^\mu+\epsilon^\mu(x)+O(\epsilon^2),
\]
condition \eqref{eq:finite-conformal-map} reduces at first order to the
conformal Killing equation \eqref{eq:conformal-killing-equation}.  If
\(f\) and \(h\) satisfy \eqref{eq:finite-conformal-map}, then their
composition satisfies it with Weyl factor
\[
  \Omega_{h\circ f}(x)=\Omega_h(f(x))\Omega_f(x).
\]
Thus finite conformal maps form a group wherever the composed maps are
defined.

Translations, rotations, and dilatations are immediate examples.  A further
basic example is inversion,
\begin{equation}
  I(x)^\mu=\frac{x^\mu}{x^2},
  \qquad
  x^2=\delta_{\mu\nu}x^\mu x^\nu .
  \label{eq:inversion-map}
\end{equation}
Its Jacobian is
\[
  \frac{\partial I^\mu}{\partial x^\nu}
  =
  \frac1{x^2}
  \left(
  \delta^\mu{}_\nu-2\frac{x^\mu x_\nu}{x^2}
  \right),
\]
and the matrix in parentheses is orthogonal.  Therefore
\[
  \Omega_I(x)=\frac1{x^2}.
\]
Inversion is conformal but lies in a disconnected component of the full
Euclidean conformal group: it cannot be reached by exponentiating a smooth
path of conformal Killing vectors from the identity.

\begin{figure}[H]
  \centering
  \resizebox{0.72\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \draw[-{Latex[length=2mm]},thick] (-4.3,0)--(4.3,0) node[right] {\(x^1\)};
    \draw[-{Latex[length=2mm]},thick] (0,-2.75)--(0,2.85) node[above] {\(x^2\)};
    \draw[thick] (0,0) circle (1.45);
    \draw[dashed] (-1.45,0)--(1.45,0);
    \node at (2.25,-1.05) {unit sphere};
    \fill (0,0) circle (0.045);
    \node[below left] at (0,0) {\(0\)};

    \coordinate (x) at (2.35,1.35);
    \coordinate (Ix) at (0.34,0.20);
    \draw[red!70!black,fill=red!70!black] (x) circle (0.055);
    \draw[green!50!black,fill=green!50!black] (Ix) circle (0.055);
    \draw[-{Latex[length=2mm]},very thick,green!50!black] (x)--(Ix);
    \node[red!70!black,right] at (x) {\(x\)};
    \node[green!50!black,below right] at (Ix) {\(I(x)=x/x^2\)};
    \node[align=center] at (0,-3.25)
      {points outside the unit sphere are sent inside, and conversely};
  \end{tikzpicture}
  }
  \caption{Euclidean inversion maps each ray through the origin to itself and
  sends the radial coordinate \(r=|x|\) to \(1/r\).}
  \label{fig:euclidean-inversion-unit-sphere}
\end{figure}

The composition
\[
  I\circ T_c\circ I
\]
with translation \(T_c(x)=x+c\) gives a special conformal
transformation
\begin{equation}
  x^\mu
  \longmapsto
  x'^\mu
  =
  \frac{x^\mu+c^\mu x^2}{1+2c\cdot x+c^2x^2}.
  \label{eq:finite-special-conformal}
\end{equation}
Expanding to first order in \(c\) gives
\[
  \delta x^\mu
  =
  c^\mu x^2-2(c\cdot x)x^\mu,
\]
which is the last term in \eqref{eq:ckv-general-solution}.
The Poincare transformations, dilatations, and inversion generate all finite
conformal transformations in \(D>2\), after restricting to domains on which
the rational maps are defined.

\section{The Conformal Algebra}

Let the quantum charges associated with the vector fields
\eqref{eq:conformal-differential-generators-1} and
\eqref{eq:conformal-differential-generators-2} be denoted by
\[
  \widehat P_\mu,\qquad
  \widehat J_{\mu\nu},\qquad
  \widehat D,\qquad
  \widehat K_\mu .
\]
Their commutators are the charge realization of the Lie bracket of conformal
Killing vector fields.  In the Euclidean convention of this chapter, indices
are contracted with \(\delta_{\mu\nu}\), and the algebra is
\begin{align}
  [\widehat D,\widehat P_\mu]&=\ii\widehat P_\mu,
  &
  [\widehat D,\widehat K_\mu]&=-\ii\widehat K_\mu,
  &
  [\widehat D,\widehat J_{\mu\nu}]&=0,
  \label{eq:conf-alg-dilatation}\\
  [\widehat P_\mu,\widehat K_\nu]
  &=
  2\ii(\delta_{\mu\nu}\widehat D-\widehat J_{\mu\nu}),
  \label{eq:conf-alg-pk}
\end{align}
together with the Poincare commutators
\begin{align}
  [\widehat J_{\mu\nu},\widehat P_\rho]
  &=
  \ii(\delta_{\rho\mu}\widehat P_\nu-\delta_{\rho\nu}\widehat P_\mu),
  \\
  [\widehat J_{\mu\nu},\widehat J_{\rho\sigma}]
  &=
  \ii(
  \delta_{\mu\rho}\widehat J_{\nu\sigma}
  -\delta_{\nu\rho}\widehat J_{\mu\sigma}
  -\delta_{\mu\sigma}\widehat J_{\nu\rho}
  +\delta_{\nu\sigma}\widehat J_{\mu\rho}).
\end{align}
All unlisted commutators among the \(\widehat P_\mu\)'s and among the
\(\widehat K_\mu\)'s vanish.

The Euclidean conformal algebra is isomorphic to \(\mathfrak{so}(D+1,1)\).
Introduce indices
\[
  A,B=-1,0,1,\ldots,D
\]
and a metric
\[
  \eta_{AB}^{(D+1,1)}=\operatorname{diag}(-1,+1,\ldots,+1),
\]
where the negative entry is the \((-1)\)-direction.  The generators
\(\widehat{\mathcal J}_{AB}=-\widehat{\mathcal J}_{BA}\) are
related to the conformal generators by
\[
  \widehat{\mathcal J}_{\mu\nu}=\widehat J_{\mu\nu},\qquad
  \widehat{\mathcal J}_{0,-1}=\widehat D,\qquad
  \widehat{\mathcal J}_{0,\mu}=\frac12(\widehat P_\mu-\widehat K_\mu),
  \qquad
  \widehat{\mathcal J}_{-1,\mu}=\frac12(\widehat P_\mu+\widehat K_\mu).
\]
Then
\[
  [\widehat{\mathcal J}_{AB},\widehat{\mathcal J}_{CD}]
  =
  \ii(
  \eta_{AC}^{(D+1,1)}\widehat{\mathcal J}_{BD}
  -\eta_{BC}^{(D+1,1)}\widehat{\mathcal J}_{AD}
  -\eta_{AD}^{(D+1,1)}\widehat{\mathcal J}_{BC}
  +\eta_{BD}^{(D+1,1)}\widehat{\mathcal J}_{AC}).
\]
The Lorentzian conformal algebra is obtained by analytic continuation of one
physical coordinate and has real form \(\mathfrak{so}(D,2)\); it is not the
real form used in the Euclidean derivation above.

\section{The Lorentzian Conformal Group and Its Cover}
\label{sec:lorentzian-conformal-group-cover}

The preceding algebra statement does not by itself identify the group that
acts on a Lorentzian Hilbert space.  Let
\[
  M=\R^{1,D-1},
  \qquad
  \eta^{\rm L}_{\mu\nu}=\operatorname{diag}(-1,+1,\ldots,+1),
  \qquad
  x^2=\eta^{\rm L}_{\mu\nu}x^\mu x^\nu .
\]
Introduce an ambient vector space
\[
  \R^{D,2}
  =
  \R X^+\oplus\R X^-\oplus \R^{1,D-1}
\]
with bilinear form
\begin{equation}
  X\cdot Y
  =
  \eta^{\rm L}_{\mu\nu}X^\mu Y^\nu
  -\frac12(X^+Y^-+X^-Y^+).
  \label{eq:ambient-d2-bilinear-form}
\end{equation}
Thus \(X\cdot X=\eta^{\rm L}_{\mu\nu}X^\mu X^\nu-X^+X^-\), and the form has
signature \((D,2)\).  The conformal compactification of Minkowski space is
the projective null cone
\begin{equation}
  \overline M
  =
  \{X\in\R^{D,2}\setminus\{0\}:X\cdot X=0\}/\R^\times .
  \label{eq:lorentzian-projective-null-cone}
\end{equation}
The Poincare patch \(X^+\ne0\) is identified with \(M\) by choosing the
representative
\begin{equation}
  X^+=1,\qquad
  X^\mu=x^\mu,\qquad
  X^-=x^2 .
  \label{eq:minkowski-patch-null-cone}
\end{equation}
The null condition follows immediately from
\eqref{eq:ambient-d2-bilinear-form}.  Points with \(X^+=0\) are the conformal
boundary of this patch.

The group \(O(D,2)\) acts linearly on \(\R^{D,2}\), preserves the null cone,
and therefore acts on \(\overline M\).  Scalar matrices act trivially after
projectivization, so the spacetime conformal transformations in the identity
component are described by the projective group
\begin{equation}
  G_{\rm conf}^{\rm Lor}
  =
  PO_0(D,2),
  \label{eq:lorentzian-projective-conformal-group}
\end{equation}
the identity component of the projective orthogonal group.  Equivalently, one
may start from \(SO_0(D,2)\) and quotient by the finite central subgroup that
acts trivially on projective null lines.  The Lie algebra of
\(G_{\rm conf}^{\rm Lor}\) is \(\mathfrak{so}(D,2)\).

The standard finite transformations on the Poincare patch are restrictions of
linear ambient transformations.  Lorentz transformations \(\Lambda\in
SO_0(1,D-1)\) act by
\[
  X^\mu\mapsto \Lambda^\mu{}_\nu X^\nu,\qquad
  X^\pm\mapsto X^\pm .
\]
Translations by \(a^\mu\) act by
\begin{equation}
  X^+\mapsto X^+,\qquad
  X^\mu\mapsto X^\mu+a^\mu X^+,\qquad
  X^-\mapsto X^-+2a\cdot X+a^2X^+,
  \label{eq:ambient-translation}
\end{equation}
where \(a\cdot X=\eta^{\rm L}_{\mu\nu}a^\mu X^\nu\).  Dilatations with
parameter \(\lambda\) act by
\begin{equation}
  X^+\mapsto e^{-\lambda}X^+,\qquad
  X^-\mapsto e^\lambda X^-,\qquad
  X^\mu\mapsto X^\mu ,
  \label{eq:ambient-dilatation}
\end{equation}
which becomes \(x^\mu\mapsto e^\lambda x^\mu\) after rescaling back to
\(X^+=1\).  Special conformal transformations with parameter \(c^\mu\) act by
\begin{equation}
  X^-\mapsto X^-,\qquad
  X^\mu\mapsto X^\mu+c^\mu X^-,\qquad
  X^+\mapsto X^+ +2c\cdot X+c^2X^- .
  \label{eq:ambient-special-conformal}
\end{equation}
On the patch \(X^+=1\), this gives
\begin{equation}
  x^\mu
  \longmapsto
  \frac{x^\mu+c^\mu x^2}{1+2c\cdot x+c^2x^2},
  \label{eq:lorentzian-finite-special-conformal}
\end{equation}
where the transformation is read locally whenever the denominator is nonzero,
and globally as a map of the compactification
\eqref{eq:lorentzian-projective-null-cone}.  These formulae show explicitly
how the infinitesimal generators integrate to the connected projective
conformal group.

For Hilbert-space representation theory, the projective spacetime group is
still not the final object.  Let
\begin{equation}
  \widetilde G_{\rm conf}^{\rm Lor}
  \longrightarrow
  G_{\rm conf}^{\rm Lor}
  \label{eq:lorentzian-conformal-universal-cover}
\end{equation}
be the universal covering group.  A unitary Lorentzian CFT carries a
strongly continuous unitary representation
\[
  U:\widetilde G_{\rm conf}^{\rm Lor}\longrightarrow
  \mathcal U(\mathcal H)
\]
whose derived representation gives the self-adjoint conformal charges on a
common invariant finite-energy domain.  The need for the cover is not a
minor global convention.  The maximal compact subalgebra of
\(\mathfrak{so}(D,2)\) is
\(\mathfrak{so}(D)\oplus\mathfrak{so}(2)\).  In the universal cover its
corresponding subgroup is covered by
\begin{equation}
  \operatorname{Spin}(D)\times \R ,
  \label{eq:maximal-compact-cover}
\end{equation}
where the \(\R\) factor is generated by Lorentzian cylinder time.  Under
radial quantization this generator is the nonnegative cylinder Hamiltonian
\(D_{\rm rad}\), and
\begin{equation}
  U_{\rm cyl}(t)=\exp(-\ii t D_{\rm rad}).
  \label{eq:cylinder-time-unitary-cover}
\end{equation}
If one forced the representation to descend to the finite group with an
\(SO(2)\) cylinder-time factor, the operator
\(\exp(-2\pi\ii D_{\rm rad})\) would have to be the identity.  This would
impose integrality of all scaling dimensions, contrary to generic unitary
CFTs.  Spinorial local operators similarly require the spin cover of spatial
rotations.  Thus later references to the Lorentzian conformal group acting on
the Hilbert space mean the universal cover
\(\widetilde G_{\rm conf}^{\rm Lor}\); its projection to
\(G_{\rm conf}^{\rm Lor}\) is the group acting on compactified spacetime
points.

The vacuum vector is invariant under
\(\widetilde G_{\rm conf}^{\rm Lor}\).  A local primary state is a lowest
energy vector for the \(\R\)-generator in
\eqref{eq:maximal-compact-cover}, transforms in a finite-dimensional unitary
representation of \(\operatorname{Spin}(D)\), and generates a positive-energy
module by the action of the noncompact translation generators.  This is the
group-theoretic meaning of the radial primary conditions used in later
chapters.

\section{Two Euclidean Dimensions}

For \(D=2\), the local conformal Killing equation has infinitely many
solutions.  In complex coordinates
\[
  z=x^1+\ii x^2,\qquad
  \bar z=x^1-\ii x^2,
\]
the local conformal transformations are
\[
  z\mapsto f(z),\qquad
  \bar z\mapsto \bar f(\bar z)
\]
in Euclidean signature, with \(f\) holomorphic in its domain.  The global
conformal transformations of the Riemann sphere are the Möbius maps
\[
  z\mapsto \frac{az+b}{cz+d},
  \qquad
  ad-bc\neq0.
\]
The infinite-dimensional enhancement has its own representation theory.  The
present core conformal block uses only the finite-dimensional global conformal
group needed for the RG fixed-point construction.
