\chapter{Unitarity Bounds and Short Multiplets}
\label{ch:volume-iii-unitarity-bounds-short-multiplets}
\ConventionsMixedCFT


\section{Positivity in Radial Quantization}

Assume the conformal theory is unitary after analytic continuation to
Lorentzian signature and satisfies reflection positivity in Euclidean radial
quantization.  Then the Hilbert space on \(S^{D-1}\) has a positive
semidefinite inner product, and null states are quotiented to obtain the
physical Hilbert space.  The adjoint relation
\[
  \widehat P_\mu^\dagger=\widehat K_\mu
\]
implies that every Gram matrix of descendant states must be positive
semidefinite.  This adjoint relation is the BPZ Hilbert-space adjoint on the
dense local finite-energy domain, derived in
\eqref{eq:bpz-p-adjoint-k-domain}; it is not a separate formal rule about
flat-space vector fields.
For example, for every complex vector \(a^\mu\) and every state created by a
local operator \(\mathcal O(0)\),
\begin{equation}
  \left\|
    a^\mu\widehat P_\mu\ket{\mathcal O}
  \right\|^2
  =
  \overline a^\mu a^\nu
  \bra{\mathcal O}\widehat K_\mu\widehat P_\nu\ket{\mathcal O}
  \ge0.
  \label{eq:first-descendant-positivity-basic}
\end{equation}
Thus \(\bra{\mathcal O}\widehat K_\mu\widehat P_\nu\ket{\mathcal O}\) is a
positive semidefinite Hermitian matrix in the vector indices and in any
additional spin or degeneracy labels.  The unitarity bounds are obtained by
computing such matrices from the conformal algebra.

In this chapter the generators are the radial-quantization generators, with
the conventional factors of \(\ii\) absorbed relative to Lorentzian Hermitian
charges.  The commutator used below is
\begin{equation}
  [\widehat K_\mu,\widehat P_\nu]
  =
  2\delta_{\mu\nu}\widehat D_{\rm rad}-2\widehat J_{\nu\mu},
  \label{eq:radial-pk-commutator}
\end{equation}
where \(\widehat J_{\mu\nu}\) acts on the spin indices of the operator
multiplet.  This convention makes the descendant Gram matrices real
Euclidean inner products.  Its sign is the radial sign derived from the
Hermitian Lorentzian charge algebra in
\eqref{eq:lorentzian-to-radial-generator-map}--\eqref{eq:radial-pk-sign-from-lorentzian};
it is not obtained by simply deleting the factors of \(\ii\) in the
Lorentzian commutators.

Let \(\ket{\mathcal O_a}\) be a primary of dimension \(\Delta\) and spin
representation \(\rho\), normalized so that
\[
  \widehat D_{\rm rad}\ket{\mathcal O_a}=\Delta\ket{\mathcal O_a},
  \qquad
  \widehat K_\mu\ket{\mathcal O_a}=0.
\]

\begin{theorem}[Level-one descendant Gram positivity]
\label{thm:level-one-descendant-gram-positivity}
Assume reflection positivity in radial quantization, the adjoint relation
\(\widehat P_\mu^\dagger=\widehat K_\mu\), and a primary multiplet
\(V_\rho=\operatorname{span}\{\ket{\mathcal O_a}\}\) of dimension
\(\Delta\).  Then the level-one descendant Gram matrix
\begin{equation}
  G_{\mu a,\nu b}^{(1)}
  =
  \bra{\mathcal O_a}\widehat K_\mu\widehat P_\nu\ket{\mathcal O_b}.
  \label{eq:level-one-gram-general}
\end{equation}
is positive semidefinite and is given by
\begin{equation}
  G_{\mu a,\nu b}^{(1)}
  =
  2
  \bra{\mathcal O_a}
  (\delta_{\mu\nu}\widehat D_{\rm rad}-\widehat J_{\nu\mu})
  \ket{\mathcal O_b}.
\end{equation}
\end{theorem}

\begin{proof}
For any finite array \(c^{\mu a}\),
\[
  \overline c^{\mu a}G^{(1)}_{\mu a,\nu b}c^{\nu b}
  =
  \left\|
    c^{\nu b}\widehat P_\nu\ket{\mathcal O_b}
  \right\|^2
  \ge0,
\]
which is the positivity assertion.  Since
\(\widehat K_\mu\ket{\mathcal O_b}=0\), the Gram matrix is
\[
  \bra{\mathcal O_a}[\widehat K_\mu,\widehat P_\nu]\ket{\mathcal O_b}.
\]
Substituting \eqref{eq:radial-pk-commutator} gives the displayed formula.
\end{proof}

This equation is a representation-theoretic constraint.  Positivity of
\(G^{(1)}\) gives lower bounds on \(\Delta\) once the \(SO(D)\) representation
\(\rho\) is specified.

The Hilbert-space convention is the following.  A primary multiplet is a
finite-dimensional subspace \(V_\rho\subset\mathcal H_{\rm rad}\) on which
\(\widehat D_{\rm rad}\) acts by the scalar \(\Delta\), and the rotation generators act
as a unitary \(SO(D)\) representation.  Descendants are obtained by applying
the closed operators \(\widehat P_\mu\) on their common finite-energy
invariant domain.  Positivity is imposed on every finite Gram matrix of
descendants in this domain; null vectors are then quotiented.  Thus each
shortening equation below is an operator equation in the quotient of the
local-operator space by null operators.

Equivalently, the calculation is the restriction of the Shapovalov form on the
conformal generalized Verma module \(M(\Delta,\rho)\) of
Definition~\ref{def:conformal-generalized-verma-module} to finite descendant
levels, with the real radial adjoint \(P_\mu^\dagger=K_\mu\).  At each level
the determinant of this form factorizes into irreducible \(SO(D)\) channels;
this finite conformal-algebra determinant is the Kac-determinant framework
used to locate singular vectors.  The proofs below compute the channels
needed here explicitly, so every stated bound follows from the displayed Gram
matrix rather than from an unstated determinant formula.

\section{Scalar Primaries}

For a scalar primary \(\phi\), the spin matrices vanish.  The level-one Gram
matrix reduces to
\[
  \bra{\phi}\widehat K_\mu\widehat P_\nu\ket{\phi}
  =
  2\Delta_\phi\delta_{\mu\nu}\braket{\phi}{\phi}.
\]
Thus
\[
  \Delta_\phi\ge0.
\]
If \(\Delta_\phi=0\), all first descendants have zero norm.  After quotienting
null states, this gives
\[
  \widehat P_\mu\ket{\phi}=0.
\]
Since \(\widehat P_\mu\) represents the derivative of the local insertion at the
origin, the corresponding local field obeys \(\partial_\mu\phi(x)=0\).  If the
theory has a unique conformally invariant vacuum and cluster decomposition
for local correlators, as specified in
Hypothesis~\ref{hyp:radial-reconstruction-data}, a spacetime-constant scalar
local field is proportional to the identity operator, so the primary belongs to
the identity representation.  These hypotheses are stated because a direct sum
with a decoupled topological or superselection sector can otherwise contain
additional zero-energy central operators.

The sharper scalar bound comes from level two.  Define
\[
  M^{(2)}_{\alpha\beta,\mu\nu}
  =
  \bra{\phi}
  \widehat K_\alpha\widehat K_\beta
  \widehat P_\mu\widehat P_\nu
  \ket{\phi}.
\]
Repeated use of the conformal commutation relations gives, for a scalar
primary,
\begin{equation}
  M^{(2)}_{\alpha\beta,\mu\nu}
  =
  4\left[
    (\delta_{\alpha\mu}\delta_{\beta\nu}
    +\delta_{\alpha\nu}\delta_{\beta\mu})
    \Delta_\phi(\Delta_\phi+1)
    -
    \delta_{\alpha\beta}\delta_{\mu\nu}\Delta_\phi
  \right]\braket{\phi}{\phi}.
  \label{eq:scalar-level-two-gram}
\end{equation}
This formula is obtained by commuting the two \(\widehat K\)'s to the right
until they annihilate the primary.  No dynamical equation has been assumed; the
only inputs are \eqref{eq:radial-pk-commutator}, the scalar action of
\(\widehat D_{\rm rad}\), and the vanishing spin action on \(\ket{\phi}\).
Contracting with \(\delta^{\alpha\beta}\delta^{\mu\nu}\) computes the norm of
\(\widehat P^2\ket{\phi}\):
\begin{equation}
  \delta^{\alpha\beta}\delta^{\mu\nu}
  M^{(2)}_{\alpha\beta,\mu\nu}
  =
  \left\|\widehat P^2\ket{\phi}\right\|^2
  =
  4\left[
    2D\,\Delta_\phi(\Delta_\phi+1)-D^2\Delta_\phi
  \right]\braket{\phi}{\phi}
  =
  4D\,\Delta_\phi\bigl(2\Delta_\phi+2-D\bigr)
  \braket{\phi}{\phi}.
  \label{eq:scalar-laplacian-descendant-norm}
\end{equation}
\begin{theorem}[Scalar unitarity bound and scalar shortening]
\label{thm:scalar-unitarity-bound-shortening}
Assume a positive-norm scalar primary \(\phi\) in a reflection-positive
radial-quantized CFT in \(D>2\).  Then either \(\Delta_\phi=0\) and
\(\widehat P_\mu\ket{\phi}=0\), or
\begin{equation}
  \Delta_\phi\ge\frac{D-2}{2}.
  \label{eq:scalar-unitarity-bound}
\end{equation}
If \(\Delta_\phi=(D-2)/2\), the scalar trace descendant is null and the
quotient local-operator space obeys
\begin{equation}
  \partial^2\phi(x)=0.
  \label{eq:scalar-shortening-laplace}
\end{equation}
\end{theorem}

\begin{proof}
The level-one Gram matrix for a scalar is
\(2\Delta_\phi\delta_{\mu\nu}\braket{\phi}{\phi}\).  Thus
\(\Delta_\phi\ge0\), and if \(\Delta_\phi=0\) then every first descendant is
null, giving \(\widehat P_\mu\ket{\phi}=0\) in the quotient Hilbert space.
For \(\Delta_\phi>0\), positivity of the level-two scalar descendant
\(\widehat P^2\ket{\phi}\) gives
\eqref{eq:scalar-laplacian-descendant-norm}; hence
\(2\Delta_\phi+2-D\ge0\), which is
\eqref{eq:scalar-unitarity-bound}.  At equality the norm in
\eqref{eq:scalar-laplacian-descendant-norm} vanishes, so the corresponding
local descendant is the Laplace equation \eqref{eq:scalar-shortening-laplace}.
\end{proof}

The \(\Delta_\phi=0\) branch was identified above as the identity
representation under the stated vacuum assumptions.  The second branch is the
scalar unitarity bound for non-identity scalar primaries.
At the level of conformal representation theory, the resulting shortened
multiplet is the massless scalar multiplet.

\section{Level-One Decomposition for Spin}

Let \(\mathcal H_\ell\) denote the \(SO(D)\) representation of rank-\(\ell\)
symmetric traceless tensors, with \(D\ge3\).  For a primary in
\(\mathcal H_\ell\), the level-one descendants transform in
\[
  \mathcal H_1\otimes\mathcal H_\ell .
\]
The \(SO(D)\)-orthogonal decomposition is
\[
  \mathcal H_1\otimes\mathcal H_\ell
  =
  \mathcal H_{\ell+1}
  \oplus
  \mathcal H_{\ell,1}
  \oplus
  \mathcal H_{\ell-1},
\]
where \(\mathcal H_{\ell,1}\) is the traceless mixed-symmetry representation
with two-row Young shape \((\ell,1)\), and absent summands are omitted.  For
\(\ell=0\) the decomposition reduces to the vector descendant of a scalar.  For
\(\ell\ge1\), the last summand is the divergence channel.

The level-one Gram operator is \(SO(D)\)-equivariant.  Because the above
decomposition is multiplicity free, Schur's lemma implies that the Gram
operator is scalar on each summand.  The scalar is computed from quadratic
Casimirs.  Write
\[
  C_\ell=\ell(\ell+D-2),
  \qquad
  C_1=D-1
\]
for the \(SO(D)\) Casimirs of \(\mathcal H_\ell\) and the vector
representation.  In the descendant space, the spin-coupling operator appearing
in \eqref{eq:level-one-gram-general} has eigenvalue
\begin{equation}
  L_Q
  =
  -\frac12\left(C_Q-C_1-C_\ell\right)
  \label{eq:spin-coupling-casimir-eigenvalue}
\end{equation}
on an irreducible summand \(Q\subset\mathcal H_1\otimes\mathcal H_\ell\).  The
minus sign is fixed by the radial commutator
\eqref{eq:radial-pk-commutator}.  The relevant Casimirs are
\begin{align}
  C_{\ell+1}
  &=
  (\ell+1)(\ell+D-1),\\
  C_{\ell,1}
  &=
  \ell(\ell+D-2)+D-3,\\
  C_{\ell-1}
  &=
  (\ell-1)(\ell+D-3).
\end{align}
Substitution in \eqref{eq:spin-coupling-casimir-eigenvalue} gives
\[
  L_{\ell+1}=-\ell,
  \qquad
  L_{\ell,1}=1,
  \qquad
  L_{\ell-1}=\ell+D-2.
\]
Since \(G^{(1)}=2(\Delta-L_Q)\) on the summand \(Q\), positivity gives
\begin{equation}
  \Delta+\ell\ge0,
  \qquad
  \Delta-1\ge0,
  \qquad
  \Delta-\ell-D+2\ge0.
  \label{eq:spin-level-one-three-inequalities}
\end{equation}
For \(\ell\ge1\), the last inequality is the strongest one.  Its null channel
is the \(\mathcal H_{\ell-1}\) projection of \(P_\mu\mathcal O\), namely the
divergence of the symmetric traceless tensor.  Thus conservation at the
unitarity bound is the null-state equation forced by positivity of the
conformal representation.

The same first-level test applies to any irreducible \(SO(D)\) representation
\(\rho\).  If \(Q\) runs over the irreducible summands in
\(\mathcal H_1\otimes\rho\), then positivity of the level-one Gram form
requires
\begin{equation}
  \Delta+\frac12\left(C_Q-C_1-C_\rho\right)\ge0
  \qquad
  \text{for every }Q\subset\mathcal H_1\otimes\rho .
  \label{eq:general-level-one-unitarity-test}
\end{equation}
When equality holds for one summand, the corresponding projected first
descendant is a null operator.  Higher-level Gram matrices can impose further
conditions in special cases, as the scalar trace descendant did above.

\section{Two-Row Mixed-Symmetry Tensor Primaries}

The first non-symmetric tensor case already displays the mechanism for
mixed-symmetry shortening.  Let \(\mathcal H_{a,b}\) be the traceless
\(SO(D)\) tensor representation with Young row lengths
\((a,b)\), where \(a\ge b\ge1\).  Work first in the stable tensor range in
which the Young diagrams displayed below are distinct irreducible tensor
representations; in lower dimensions Hodge duality may split or identify
summands, and the same Gram-matrix test is then applied to the resulting
irreducible \(\operatorname{Spin}(D)\) summands.

With the normalization \(C_1=D-1\) for the vector representation, the
quadratic Casimir of a traceless tensor representation with row lengths
\(\lambda=(\lambda_1,\lambda_2,\ldots)\) is
\begin{equation}
  C_\lambda
  =
  \sum_{i\ge1}\lambda_i(\lambda_i+D-2i).
  \label{eq:soD-tensor-casimir-row-lengths}
\end{equation}
To derive this formula, let \(r=\lfloor D/2\rfloor\) and choose the
orthonormal basis \(e_i\) for the weight space of
\(\mathfrak{so}(D)_\mathbb C\) in which the vector representation has highest
weight \(e_1\).  In the stable tensor range the highest weight of the tensor
representation is \(\Lambda=\sum_i\lambda_i e_i\), while the Weyl vector
\(\varrho\) obeys \(2\varrho_i=D-2i\).  The quadratic Casimir eigenvalue is
\((\Lambda,\Lambda+2\varrho)\), normalized so that the vector highest weight
\(e_1\) has \(C_1=(e_1,e_1+2\varrho)=D-1\), and this gives
\eqref{eq:soD-tensor-casimir-row-lengths}.
Thus
\[
  C_{a,b}=a(a+D-2)+b(b+D-4).
\]
The level-one descendant space decomposes by adding or removing one box from
the Young diagram.  The derivative vector index may be projected into an
added first, second, or third row, or it may be contracted with an index in
the first or second row and then projected back to the traceless Young
symmetry.  This gives
\begin{align}
  \mathcal H_1\otimes\mathcal H_{a,b}
  &=
  \mathcal H_{a+1,b}
  \oplus
  \mathcal H_{a,b+1}
  \oplus
  \mathcal H_{a,b,1}
  \oplus
  \mathcal H_{a-1,b}
  \oplus
  \mathcal H_{a,b-1}.
  \label{eq:two-row-vector-tensor-decomposition}
\end{align}
Invalid summands are omitted: for example \(\mathcal H_{a,b+1}\) and
\(\mathcal H_{a-1,b}\) are absent when \(a=b\), and
\(\mathcal H_{a,b-1}\) becomes \(\mathcal H_a\) when \(b=1\).  The
\(\mathcal H_{a,b,1}\) channel is present only when the corresponding
three-row tensor representation is not removed by a low-dimensional duality.
Each displayed summand has multiplicity one in the stable range, so the
level-one Gram operator is diagonal on these channels.

Substituting the Casimirs in
\eqref{eq:general-level-one-unitarity-test} gives the following table:
\begin{equation}
\begin{array}{c|c|c}
  Q
  &
  \frac12(C_Q-C_1-C_{a,b})
  &
  \text{level-one positivity condition}
  \\ \hline
  \mathcal H_{a+1,b}
  &
  a
  &
  \Delta+a\ge0
  \\
  \mathcal H_{a,b+1}
  &
  b-1
  &
  \Delta+b-1\ge0
  \\
  \mathcal H_{a,b,1}
  &
  -2
  &
  \Delta-2\ge0
  \\
  \mathcal H_{a-1,b}
  &
  -(a+D-2)
  &
  \Delta-a-D+2\ge0
  \\
  \mathcal H_{a,b-1}
  &
  -(b+D-3)
  &
  \Delta-b-D+3\ge0 .
\end{array}
\label{eq:two-row-level-one-bound-table}
\end{equation}
For instance,
\[
  \frac12(C_{a-1,b}-C_1-C_{a,b})
  =
  -a-D+2,
\]
because replacing \(a\) by \(a-1\) changes the first-row contribution to the
Casimir by \(-2a-D+3\), and the vector Casimir subtracts another \(D-1\).
The other entries follow from the same row-by-row subtraction in
\eqref{eq:soD-tensor-casimir-row-lengths}.

\begin{proposition}[Two-row tensor unitarity bound at first level]
\label{prop:two-row-mixed-symmetry-unitarity-bound}
Let \(\mathcal O\) be a primary whose rotation representation is
\(\mathcal H_{a,b}\), \(a\ge b\ge1\), in the stable tensor range described
above.  If \(a>b\), positivity of the first descendant Gram matrix implies
\begin{equation}
  \Delta_{\mathcal O}\ge a+D-2.
  \label{eq:two-row-nonrectangular-bound}
\end{equation}
At equality the \(\mathcal H_{a-1,b}\) projection of
\(P_\mu|\mathcal O\rangle\) is null.  In local-operator notation this is
\begin{equation}
  \Pi_{a-1,b}\,\partial\mathcal O=0,
  \label{eq:two-row-nonrectangular-shortening}
\end{equation}
where \(\Pi_{a-1,b}:\mathcal H_1\otimes\mathcal H_{a,b}\to
\mathcal H_{a-1,b}\) is the \(SO(D)\)-equivariant traceless Young projector.
If \(a=b\), the channel \(\mathcal H_{a-1,b}\) is not an allowed Young shape,
and the leading first-level condition is instead
\begin{equation}
  \Delta_{\mathcal O}\ge a+D-3.
  \label{eq:two-row-rectangular-bound}
\end{equation}
At equality the \(\mathcal H_{a,a-1}\) projection of the first descendant is
null:
\begin{equation}
  \Pi_{a,a-1}\,\partial\mathcal O=0.
  \label{eq:two-row-rectangular-shortening}
\end{equation}
\end{proposition}

\begin{proof}
The inequalities in \eqref{eq:two-row-level-one-bound-table} are precisely
the eigenvalue inequalities for the positive semidefinite level-one Gram form.
When \(a>b\), the removal channel \(\mathcal H_{a-1,b}\) is present, and its
condition \(\Delta\ge a+D-2\) is stronger than
\(\Delta\ge b+D-3\) because \(a-b+1>0\); it is also stronger than the
add-a-box conditions for a positive-energy representation.  If equality holds,
the Gram norm on that irreducible summand is zero, hence the corresponding
state is in the null subspace and becomes the operator equation
\eqref{eq:two-row-nonrectangular-shortening} in the quotient by null
operators.

When \(a=b\), the Young diagram \((a-1,a)\) is not allowed, so the
\(\mathcal H_{a-1,b}\) row in the table is absent.  The strongest remaining
condition is \(\Delta\ge a+D-3\), coming from
\(\mathcal H_{a,a-1}\).  Equality makes the
\(\mathcal H_{a,a-1}\) component of \(P_\mu|\mathcal O\rangle\) null, giving
\eqref{eq:two-row-rectangular-shortening}.  The special case
\((a,b)=(1,1)\) is the antisymmetric two-tensor primary: the shortening
equation is the projected divergence \(\partial^\mu J_{\mu\nu}=0\), after
traces are removed according to the representation.
\end{proof}

\section{Spinor Primaries}

For spinor primaries the rotation representation is a representation of
\(\operatorname{Spin}(D)\), the double cover of \(SO(D)\).  Let \(D\ge3\), and
let \(S\) be an irreducible complex spinor representation.  In even \(D\),
\(S\) may be one of the two chiral spinor representations; Clifford
multiplication by a vector then maps \(S\) to the opposite chirality.  Denote
the spinor representation reached by Clifford multiplication by \(S_\gamma\).
Thus \(S_\gamma\simeq S\) in odd dimension, while \(S_\gamma\) is the opposite
chiral spinor in even dimension.  Choose Euclidean Clifford maps
\[
  \Gamma_\mu:S\longrightarrow S_\gamma,
  \qquad
  \Gamma_\mu^\dagger:S_\gamma\longrightarrow S,
  \qquad
  \Gamma_\mu^\dagger\Gamma_\nu+\Gamma_\nu^\dagger\Gamma_\mu
  =
  2\delta_{\mu\nu}\id_S .
\]
These maps are intertwiners for the \(\operatorname{Spin}(D)\) action and
therefore identify the gamma-trace summand in the level-one descendant space.

\begin{theorem}[Spinor unitarity bound and Dirac shortening]
\label{thm:spinor-unitarity-bound}
Let \(\psi\) be a spinor primary of scaling dimension \(\Delta_\psi\) whose
radial-quantization primary space carries the unitary spin representation
\(S\).  Then
\begin{equation}
  \Delta_\psi\ge\frac{D-1}{2}.
  \label{eq:spinor-primary-unitarity-bound}
\end{equation}
If equality holds, the gamma-trace first descendant is null.  In local-operator
notation the shortening equation is
\begin{equation}
  \Gamma^\mu\partial_\mu\psi(x)=0 .
  \label{eq:spinor-dirac-shortening}
\end{equation}
\end{theorem}

\begin{proof}
The relevant \(SO(D)\) Casimir eigenvalues, in the normalization used in
\eqref{eq:spin-coupling-casimir-eigenvalue}, are
\[
  C_S=C_{S_\gamma}=\frac{D(D-1)}8,
  \qquad
  C_1=D-1 .
\]
The tensor product of the vector representation with \(S\) decomposes as
\[
  \mathcal H_1\otimes S
  =
  S_\gamma\oplus S_{3/2},
\]
where \(S_{3/2}\) is the gamma-traceless vector-spinor representation.  The
orthogonal projection onto the first summand is the gamma-trace projection:
for \(A_\mu\in\mathcal H_1\otimes S\),
\[
  (\Pi_\gamma A)_\mu
  =
  \frac1D\,\Gamma_\mu^\dagger\Gamma^\nu A_\nu,
  \qquad
  \Gamma^\mu(A_\mu-\Pi_\gamma A_\mu)=0 .
\]
Equivalently, after normalizing the primary state so that
\(\|\ket{\psi(s)}\|^2=\|s\|_S^2\) for \(s\in S\), the gamma-trace descendant
has Gram norm
\begin{equation}
  \left\|
    \frac1{\sqrt D}\Gamma^\mu\widehat P_\mu\ket{\psi(s)}
  \right\|^2
  =
  2\left(\Delta_\psi-\frac{D-1}{2}\right)\|s\|_S^2 .
  \label{eq:spinor-gamma-trace-descendant-norm}
\end{equation}
This is precisely \eqref{eq:level-one-gram-general} restricted to the
\(Q=S_\gamma\) irreducible summand, because
\[
  \Delta_\psi+\frac12(C_{S_\gamma}-C_1-C_S)
  =
  \Delta_\psi-\frac{D-1}{2}.
\]
Reflection positivity of radial quantization requires the left-hand side of
\eqref{eq:spinor-gamma-trace-descendant-norm} to be nonnegative for every
\(s\), giving \eqref{eq:spinor-primary-unitarity-bound}.  If equality holds,
the gamma-trace descendant is a null state and therefore vanishes in the
quotient local-operator space, which gives
\eqref{eq:spinor-dirac-shortening}.

For completeness, the complementary vector-spinor summand has
\[
  C_{S_{3/2}}=C_S+D .
\]
Its level-one inequality is
\[
  \Delta_\psi+\frac12(C_{S_{3/2}}-C_1-C_S)
  =
  \Delta_\psi+\frac12\ge0,
\]
so it gives no stronger lower bound for a positive-energy spinor primary.
\end{proof}

\section{Symmetric Traceless Tensor Primaries}

Let
\[
  \mathcal O_{\mu_1\cdots\mu_\ell}(x)
\]
be a primary transforming as a rank-\(\ell\) symmetric traceless tensor of
\(SO(D)\), with \(\ell\ge1\).  Thus the indices are symmetric and
\[
  \delta^{\mu_1\mu_2}
  \mathcal O_{\mu_1\mu_2\cdots\mu_\ell}=0.
\]
\begin{theorem}[Symmetric-traceless unitarity bound]
\label{thm:stt-unitarity-bound-shortening}
Assume a positive-norm primary
\(\mathcal O_{\mu_1\cdots\mu_\ell}\), \(\ell\ge1\), transforming in the
rank-\(\ell\) symmetric traceless representation of \(SO(D)\), with \(D\ge3\).
Then
\begin{equation}
  \Delta_{\mathcal O}\ge \ell+D-2.
  \label{eq:spin-ell-unitarity-bound}
\end{equation}
At equality, the divergence descendant is null and therefore
\begin{equation}
  \partial^{\mu_1}
  \mathcal O_{\mu_1\mu_2\cdots\mu_\ell}(x)=0.
  \label{eq:spin-ell-conservation-shortening}
\end{equation}
\end{theorem}

\begin{proof}
The level-one descendant space decomposes as
\[
  \mathcal H_1\otimes\mathcal H_\ell
  =
  \mathcal H_{\ell+1}\oplus\mathcal H_{\ell,1}\oplus\mathcal H_{\ell-1}.
\]
By Theorem~\ref{thm:level-one-descendant-gram-positivity}, the Gram operator is
positive on each irreducible summand.  The Casimir computation above gives the
three inequalities in \eqref{eq:spin-level-one-three-inequalities}; for
\(\ell\ge1\), the strongest is
\(\Delta_{\mathcal O}-\ell-D+2\ge0\).  The corresponding summand is
\(\mathcal H_{\ell-1}\), obtained by taking the divergence of the symmetric
traceless tensor.  Saturation therefore makes precisely the divergence
descendant null, giving \eqref{eq:spin-ell-conservation-shortening} in the
quotient local-operator space.
\end{proof}

For \(\ell=1\) this is current conservation and
\(\Delta_J=D-1\).  For \(\ell=2\) it is stress-tensor conservation and
\(\Delta_T=D\).

\section{Short Multiplets as Operator Equations}

\begin{theorem}[Basic unitarity bounds and short multiplets]
\label{thm:basic-unitarity-bounds-short-multiplets}
Let \(D\ge3\), and assume reflection-positive radial quantization, the adjoint
relation \(P_\mu^\dagger=K_\mu\), and the quotient of null states in the local
operator space.  Then the following bounds and equality conditions hold for
positive-norm primaries.
\begin{enumerate}
  \item A scalar primary \(\phi\) either lies in the \(\Delta_\phi=0\)
  identity branch described in
  Theorem~\ref{thm:scalar-unitarity-bound-shortening}, or obeys
  \[
    \Delta_\phi\ge\frac{D-2}{2}.
  \]
  At \(\Delta_\phi=(D-2)/2\), the level-two trace descendant is null and
  \[
    \partial^2\phi=0
  \]
  in the quotient local-operator space.
  \item A spinor primary \(\psi\) obeys
  \[
    \Delta_\psi\ge\frac{D-1}{2}.
  \]
  At equality, the gamma-trace first descendant is null and
  \[
    \Gamma^\mu\partial_\mu\psi=0.
  \]
  \item A rank-\(\ell\) symmetric traceless tensor primary
  \(\mathcal O_{\mu_1\cdots\mu_\ell}\), \(\ell\ge1\), obeys
  \[
    \Delta_{\mathcal O}\ge\ell+D-2.
  \]
  At equality, the divergence first descendant is null and
  \[
    \partial^{\mu_1}\mathcal O_{\mu_1\mu_2\cdots\mu_\ell}=0.
  \]
\end{enumerate}
\end{theorem}

\begin{proof}
The scalar statement is Theorem~\ref{thm:scalar-unitarity-bound-shortening},
which uses the level-one scalar Gram matrix and the level-two trace-descendant
norm \eqref{eq:scalar-laplacian-descendant-norm}.  The spinor statement is
Theorem~\ref{thm:spinor-unitarity-bound}, obtained from the gamma-trace
summand of the level-one descendant Gram matrix.  The symmetric-traceless
statement is Theorem~\ref{thm:stt-unitarity-bound-shortening}, obtained from
the divergence channel \(\mathcal H_{\ell-1}\) in
\(\mathcal H_1\otimes\mathcal H_\ell\).  In all three cases, equality means
that the displayed descendant has zero radial norm.  Reflection positivity
therefore puts that descendant in the null radical, and the null quotient
turns it into the corresponding local operator equation.
\end{proof}

\section{Short Multiplets as Operator Equations}
\label{sec:short-multiplets-operator-equations}

A short conformal multiplet is a primary multiplet for which a descendant has
zero norm and is set to zero in the physical Hilbert space.  In local-operator
language, the null descendant is an operator equation.  The equations
\[
  \partial^2\phi=0,\qquad
  \partial^\mu J_\mu=0,\qquad
  \partial^\mu T_{\mu\nu}=0
\]
are therefore representation-theoretic statements in a unitary conformal
theory.  They also carry dynamical information: a conserved current generates
a global symmetry, the stress tensor generates spacetime symmetries, and a
scalar saturating the scalar bound obeys the scalar Laplace shortening equation.

The unitarity bounds thus convert positivity of the Hilbert-space inner
product into exact restrictions on the allowed local operator spectrum.
