\chapter{Fixed Points and Conformal Data}
\label{ch:volume-iii-fixed-points-conformal-data}
\ConventionsMixedCFT


\section{Local QFT Data Used in Fixed-Point Analysis}

Let \(D\geq2\) be the spacetime dimension, let \(d=D-1\), and let
\(\eta_{\mu\nu}=\operatorname{diag}(-,+,\ldots,+)\).  The local Lorentzian
data used in this volume are the following.
\begin{enumerate}
  \item A Hilbert space \(\Hilb\) of physical states.
  \item A strongly continuous unitary representation \(U(\Lambda,a)\) of the
  proper orthochronous Poincare group, or of its spin cover when spinorial
  fields occur.  The group law is
  \[
    U(\Lambda',a')U(\Lambda,a)
    =
    U(\Lambda'\Lambda,\Lambda'a+a') .
  \]
  \item Self-adjoint generators \(P_\mu\) and \(J_{\mu\nu}=-J_{\nu\mu}\),
  normalized near the identity by
  \[
    U(\Lambda,a)
    =
    1+\ii a^\mu P_\mu
    +\frac{\ii}{2}\omega^{\mu\nu}J_{\mu\nu}
    +O(a^2,\omega^2,a\omega),
    \qquad
    \Lambda^\mu{}_\nu=\delta^\mu{}_\nu+\omega^\mu{}_\nu .
  \]
  The operators \(P_\mu\) generate energy and momentum, while \(J_{\mu\nu}\)
  generates spatial rotations and Lorentz boosts.
  \item A Poincare-invariant unit vector \(\vac\in\Hilb\), so
  \(U(\Lambda,a)\vac=\vac\).
  \item Local fields \(\widehat\Phi_\alpha(x)\), understood as
  operator-valued distributions on a common invariant dense domain, with
  Poincare covariance
  \[
    U(\Lambda,a)^{-1}\widehat\Phi_\alpha(x)U(\Lambda,a)
    =
    R(\Lambda)_\alpha{}^\beta\,
    \widehat\Phi_\beta(\Lambda x+a).
  \]
  Here \(R\) is the finite-dimensional Lorentz representation carried by the
  field component label \(\alpha\).  Bosonic local fields commute, and
  fermionic local fields anticommute, after smearing with test functions whose
  supports are spacelike separated.
\end{enumerate}

In addition, the discussion of conformal symmetry assumes a renormalized
local stress tensor \(\widehat T_{\mu\nu}(x)\).  It is a symmetric tensor-valued
operator distribution satisfying
\begin{equation}
  \partial_\mu\widehat T^{\mu\nu}=0,
  \qquad
  \widehat T^{\mu\nu}=\widehat T^{\nu\mu}.
  \label{eq:cft-opening-stress-conservation}
\end{equation}
where the conservation equation is distributional.  When the spatial
integrals exist with the chosen domain and infrared prescription, the
translation and Lorentz generators are represented by
\begin{equation}
  P^\mu
  =
  \int_{\Sigma_t}\dd^d\vec x\,
  \widehat T^{0\mu}(t,\vec x).
  \label{eq:cft-opening-p-charge}
\end{equation}
and
\begin{equation}
  J^{\mu\nu}
  =
  \int_{\Sigma_t}\dd^d\vec x\,
  \bigl(
    x^\mu\widehat T^{0\nu}(t,\vec x)
    -
    x^\nu\widehat T^{0\mu}(t,\vec x)
  \bigr).
  \label{eq:cft-opening-j-charge}
\end{equation}
with \(\Sigma_t=\{x^0=t\}\).  These equations are operator equations after
the same smearing, renormalization, and boundary conventions used to define
\(\widehat T_{\mu\nu}\).

Lorentz covariance of the stress tensor means
\begin{equation}
  U(\Lambda,0)\widehat T_{\mu\nu}(x)U(\Lambda,0)^{-1}
  =
  \Lambda_\mu{}^\rho\Lambda_\nu{}^\sigma
  \widehat T_{\rho\sigma}(\Lambda x).
  \label{eq:cft-opening-stress-lorentz}
\end{equation}
where indices on \(\Lambda\) are moved with \(\eta_{\mu\nu}\).  Hence
\(\widehat T^\mu{}_\mu(x)\) is a scalar local operator:
\[
  U(\Lambda,0)\widehat T^\mu{}_\mu(x)U(\Lambda,0)^{-1}
  =
  \widehat T^\mu{}_\mu(\Lambda x).
\]

\begin{remark}[Fixed points as supplied data, not universal ultraviolet origins]
\label{rem:cft-volume-perturbation-scope}
This volume analyzes the consequences of fixed-point data once those data have
been constructed or assumed.  A conformal perturbation problem starts with a
specific fixed point \(\mathcal C\), a specified family of scalar local
operators \(O_A\), and a prescription for the contact terms in products of
\(\int O_A\).  The deformation coordinates \(g^A\) then describe a local
chart near \(\mathcal C\), not a general definition of QFT.  In particular,
the existence of an ultraviolet-complete theory does not follow from, nor is
it equivalent to, being representable as a relevant deformation of an
accessible CFT.  Asymptotically free gauge theories are controlled in the
ultraviolet by running local couplings and require a nonperturbative
definition such as a lattice or other regulator framework; constructive and
SPDE routes provide different fixed-point or continuum-limit data.  The
conformal methods developed below are used precisely where their hypotheses
are present.
\end{remark}

\section{Renormalized Operators and the Trace Equation}

Fix a renormalized coordinate chart \(\{g_I(\mu)\}_{I\in\mathcal I}\) on a
space of local scalar deformations.  The index set \(\mathcal I\) labels
renormalized local insertions \([O_I]_\mu(x)\).  In a Lagrangian presentation
with bare scalar operators \(O_I^{\rm bare}\), write the interaction density as
\begin{equation}
  \mathcal L_{\rm int}^{\rm bare}
  =
  \sum_I g_I^{\rm bare}O_I^{\rm bare}.
  \label{eq:cft-opening-bare-interaction}
\end{equation}
The renormalized insertions are defined by the response of the bare density to
renormalized coordinates:
\begin{equation}
  \delta\mathcal L_{\rm int}^{\rm bare}
  =
  \sum_I \delta g_I^{\rm bare}O_I^{\rm bare}
  =
  \sum_I \delta g_I(\mu)\,[O_I]_\mu .
  \label{eq:cft-opening-renormalized-insertions}
\end{equation}
This equation fixes the normalization and mixing convention for the local
operators in the chosen scheme.  The beta-function vector field is
\begin{equation}
  \frac{\dd}{\dd\log\mu}g_I(\mu)
  =
  \beta_I(g(\mu)).
  \label{eq:cft-opening-beta-function}
\end{equation}

\begin{proposition}[Flat trace equation from a renormalized source chart]
\label{prop:cft-opening-flat-trace-equation}
Assume that the renormalized chart
\(\{g_I(\mu),[O_I]_\mu\}_{I\in\mathcal I}\) is supplied by a local source
functional whose constant Weyl variation is governed, in flat space, by the
local RG vector field
\[
  \Delta_\sigma^{\rm loc}
  =
  \int\dd^D x\,\sigma
  \left(
    2g_{\mu\nu}\frac{\delta}{\delta g_{\mu\nu}}
    +
    \beta_I(g)\frac{\delta}{\delta g_I}
  \right)
\]
up to local anomaly/contact functionals and up to the divergence of a local
virial current.  Then, as a renormalized flat-space local-operator identity
away from coincident insertions,
\begin{equation}
  \widehat T^\mu{}_\mu
  =
  \sum_I\beta_I(g)\,[O_I]_\mu
  +
  \partial_\mu V^\mu .
  \label{eq:cft-opening-trace-beta}
\end{equation}
The vector \(V^\mu\) is a local vector operator recording the virial and
improvement ambiguity in the chosen stress-tensor representative.  Contact
terms in \eqref{eq:cft-opening-trace-beta} are not inferred from
separated-point kernels; they are fixed together with the renormalized source
Ward identities of the chart.  Curvature terms appear after coupling the
theory to a background metric, but vanish in the flat separated-point identity
displayed here.
\end{proposition}

\begin{proof}
The source chart gives two descriptions of the same infinitesimal change of
renormalization length.  First, varying the metric by a constant Weyl factor
\(\delta_\sigma g_{\mu\nu}=2\sigma g_{\mu\nu}\) gives, by the stress-tensor
definition,
\[
  \delta_\sigma W
  =
  -\int\dd^D x\,\sqrt g\,\sigma\,
  \left\langle T^\mu{}_\mu(x)\right\rangle .
\]
Second, the same change expressed at fixed regulator is a displacement along
the beta-function vector field in the renormalized coordinates:
\[
  \delta_\sigma W
  =
  -\int\dd^D x\,\sqrt g\,\sigma\,
  \beta_I(g)\left\langle [O_I]_\mu(x)\right\rangle
  \]
with the sign fixed by the convention
\(\mu\,\dd g_I/\dd\mu=\beta_I\) and by
\eqref{eq:cft-opening-renormalized-insertions}.  Equating the two
descriptions, and restoring the possible local virial term produced by the
choice of stress-tensor representative, gives
\eqref{eq:cft-opening-trace-beta} inside correlators at separated points.
Local anomaly terms and derivatives of delta functions are exactly the contact
terms belonging to the chosen source chart, so they do not modify the
displayed separated-point flat-space operator identity.
\end{proof}

\begin{remark}[Scale invariance versus conformal invariance]
\label{rem:cft-opening-scale-versus-conformal}
At a fixed point the vanishing of the beta-function part of
\eqref{eq:cft-opening-trace-beta} gives
\begin{equation}
  \widehat T^\mu{}_\mu=\partial_\mu V^\mu
  \label{eq:cft-opening-scale-trace-virial}
\end{equation}
in the flat-space local-operator quotient determined by the source chart.  If
\(\widehat T^{\mu\nu}\) is symmetric and conserved, then
\begin{equation}
  D^\mu=x_\nu\widehat T^{\mu\nu}-V^\mu
  \label{eq:cft-opening-dilatation-current}
\end{equation}
is a conserved dilatation current exactly when
\eqref{eq:cft-opening-scale-trace-virial} holds.  Conformal symmetry requires
the stronger local statement that the virial class is removable by the allowed
stress-tensor improvements.  For the scalar improvement
\[
  \widehat T^{\mu\nu}\longmapsto
  \widehat T^{\mu\nu}
  +
  (\partial^\mu\partial^\nu-\eta^{\mu\nu}\partial^2)L ,
\]
the trace changes by \(-(D-1)\partial^2L\), so the trace can be removed when
\begin{equation}
  V^\mu=(D-1)\partial^\mu L+J^\mu,
  \qquad
  \partial_\mu J^\mu=0,
  \label{eq:cft-opening-removable-virial}
\end{equation}
modulo the operator identities and contact-term convention of the source chart.
More general improvement tensors enlarge the same quotient by the image of
their trace-changing map.  Thus the scale-to-conformal implication is the
local problem of proving that this virial obstruction class vanishes.

In this stress-tensor framework, a scale-invariant but nonconformal local QFT
is a theory for which \eqref{eq:cft-opening-scale-trace-virial} gives a
conserved dilatation current \(D^\mu=x_\nu\widehat T^{\mu\nu}-V^\mu\), while
the class of \(V^\mu\) is nonzero in the quotient by conserved currents and
allowed improvement currents.  Equivalently, the theory has a virial current
of dimension \(D-1\) whose divergence is the trace, but no admissible
improvement makes the stress tensor traceless.  This is the obstruction
isolated by Polchinski's analysis of scale versus conformal invariance:
\(D^\mu=x_\nu T^{\mu\nu}-V^\mu\) may generate scale transformations even when
there are no conserved special-conformal currents.\footnote{J.~Polchinski,
``Scale and conformal invariance in quantum field theory,'' \emph{Nucl. Phys.
B} \textbf{303} (1988) 226--236.}

The status is dimension-dependent.  In \(D=2\), the
Zamolodchikov--Polchinski argument proves the implication under explicit
Hilbert-space positivity, locality, stress-tensor, vacuum, spectrum, and
regularity hypotheses strong enough for the \(c\)-function and stress-tensor
two-point analysis.  In \(D=4\), the Luty--Polchinski--Rattazzi program and
related anomaly/dilaton arguments give powerful conditional constraints, but
this monograph does not use an unconditional theorem asserting that every
unitary scale-invariant four-dimensional QFT is conformal.  In \(D=3\) and
\(D>4\), later uses of conformal symmetry likewise require the traceless
improved stress tensor, the vanishing virial class, or the equivalent conformal
Ward identities as part of the stated data.  The cohomological formulation and
references are given in
Remark~\ref{rem:scale-conformal-dimension-status}.
\end{remark}

\begin{proposition}[Virial class, improvement, and conformal currents]
\label{prop:cft-opening-virial-improvement-currents}
Let \(T^{\mu\nu}\) be a symmetric conserved stress tensor in flat
\(D\)-dimensional Minkowski space.  Suppose that
\[
  T^\mu{}_\mu=\partial_\mu V^\mu
\]
as a local operator identity in the chosen separated-point/contact chart.
Then
\[
  D^\mu=x_\nu T^{\mu\nu}-V^\mu
\]
is a conserved dilatation current.  If, for some scalar local operator \(L\)
and conserved current \(J^\mu\),
\[
  V^\mu=(D-1)\partial^\mu L+J^\mu,
  \qquad
  \partial_\mu J^\mu=0,
\]
then the improved stress tensor
\[
  T_{\rm imp}^{\mu\nu}
  =
  T^{\mu\nu}
  +
  (\partial^\mu\partial^\nu-\eta^{\mu\nu}\partial^2)L
\]
is symmetric, conserved, and traceless.  Consequently, for every conformal
Killing vector \(\xi^\mu\) satisfying
\[
  \partial_\mu\xi_\nu+\partial_\nu\xi_\mu
  =
  \frac2D(\partial_\rho\xi^\rho)\eta_{\mu\nu},
\]
the current \(j_\xi^\mu=T_{\rm imp}^{\mu\nu}\xi_\nu\) is conserved.
\end{proposition}

\begin{proof}
The divergence of the proposed dilatation current is
\[
  \partial_\mu D^\mu
  =
  T^\mu{}_\mu
  +x_\nu\partial_\mu T^{\mu\nu}
  -\partial_\mu V^\mu
  =
  0 .
\]
The improvement term is symmetric.  Its divergence vanishes because ordinary
derivatives commute:
\[
  \partial_\mu
  (\partial^\mu\partial^\nu-\eta^{\mu\nu}\partial^2)L
  =
  \partial^\nu\partial^2L-\partial^\nu\partial^2L
  =
  0 .
\]
Its trace is
\[
  \eta_{\mu\nu}
  (\partial^\mu\partial^\nu-\eta^{\mu\nu}\partial^2)L
  =
  -(D-1)\partial^2L .
\]
Therefore
\[
  T_{{\rm imp}\,\mu}^{\ \ \ \mu}
  =
  \partial_\mu V^\mu-(D-1)\partial^2L
  =
  \partial_\mu J^\mu
  =
  0 .
\]
For a conformal Killing vector,
\[
  \partial_\mu j_\xi^\mu
  =
  T_{\rm imp}^{\mu\nu}\partial_\mu\xi_\nu
  =
  \frac12T_{\rm imp}^{\mu\nu}
  (\partial_\mu\xi_\nu+\partial_\nu\xi_\mu)
  =
  \frac1D(\partial_\rho\xi^\rho)
  T_{{\rm imp}\,\mu}^{\ \ \ \mu}
  =
  0 ,
\]
where conservation and symmetry of \(T_{\rm imp}^{\mu\nu}\) were used.  Thus
the improved traceless representative supplies the currents for translations,
rotations, dilatations, and special conformal transformations whenever the
corresponding smeared charges are defined.
\end{proof}

An RG fixed point in the coordinate chart is a point \(g_\ast\) satisfying
\begin{equation}
  \beta_I(g_\ast)=0
  \qquad
  \text{for every }I\in\mathcal I .
  \label{eq:cft-opening-fixed-point}
\end{equation}
At such a point the beta-function part of the trace vanishes.  A flat-space
conformal fixed point is obtained when the stress tensor can be chosen, by
allowed improvements and contact-term conventions, so that
\begin{equation}
  \widehat T^\mu{}_\mu=0
  \label{eq:cft-opening-traceless-stress}
\end{equation}
as a renormalized local operator equation in flat space.  Equation
\eqref{eq:cft-opening-traceless-stress}, together with
\eqref{eq:cft-opening-stress-conservation}, is the local input from which the
conformal currents and charges are derived in the next chapter.

The linearized RG flow near \(g_\ast\) records the scaling dimensions of
deformations.  Let
\[
  B_{IJ}
  =
  \left.
  \frac{\partial\beta_I}{\partial g_J}
  \right|_{g=g_\ast}.
\]
Choose an eigenvector \(v^{(a)}_J\) and write
\[
  g_I-g_{\ast I}=\lambda_a v^{(a)}_I+O(\lambda^2),
  \qquad
  \sum_J B_{IJ}v^{(a)}_J=-y_a v^{(a)}_I .
\]
With the momentum scale \(\mu\) as flow parameter,
\[
  \mu\frac{\dd\lambda_a}{\dd\mu}
  =
  -y_a\lambda_a+O(\lambda^2).
\]
Equivalently, under a length rescaling \(x\mapsto b x\), \(b>1\),
\[
  \lambda_a(b)=b^{y_a}\lambda_a+O(\lambda^2).
\]
If the corresponding scalar perturbation is
\(\int\dd^Dx\,\lambda_a O_a(x)\), then
\begin{equation}
  y_a=D-\Delta_a,
  \label{eq:cft-opening-y-delta}
\end{equation}
where \(\Delta_a\) is the scaling dimension of \(O_a\).  Thus \(y_a>0\),
\(y_a=0\), and \(y_a<0\) define relevant, marginal, and irrelevant
directions, respectively, in the long-distance flow.

\begin{proposition}[Linearized RG eigenvalues and operator dimensions]
\label{prop:cft-opening-linearized-rg-dimensions}
Assume that the beta-function vector field is differentiable at \(g_\ast\)
and that the tangent direction \(v^{(a)}\) is an ordinary eigenvector of
\(B_{IJ}=\partial\beta_I/\partial g_J|_{g_\ast}\) with eigenvalue \(-y_a\).
Let
\[
  O_a=\sum_I v^{(a)}_I[O_I]_{\ast}
\]
be the corresponding scalar deformation operator in the linearized source
chart.  If \(O_a\) is a scaling operator at the fixed point, then its scaling
dimension satisfies
\[
  \Delta_a=D-y_a .
\]
If \(B\) has a Jordan block instead of a diagonal eigenspace, the corresponding
operators form a logarithmic mixing block under scale transformations; the
number \(D-y_a\) is then the common diagonal scaling dimension of the block,
not a complete diagonalization of the dilatation action.
\end{proposition}

\begin{proof}
Along the eigen-direction write
\[
  g_I(\mu)-g_{\ast I}=\lambda_a(\mu)v^{(a)}_I+O(\lambda_a^2).
\]
The RG equation gives
\[
  \mu\frac{\dd\lambda_a}{\dd\mu}
  =
  -y_a\lambda_a+O(\lambda_a^2).
\]
Changing the unit of length by \(x\mapsto bx\), \(b>1\), corresponds to
lowering the momentum scale by \(b^{-1}\), hence
\[
  \lambda_a(b)=b^{y_a}\lambda_a+O(\lambda_a^2).
\]
At the fixed point a scalar scaling operator obeys
\[
  O_a(x)\mapsto b^{-\Delta_a}O_a(b^{-1}x).
\]
Therefore the integrated perturbation transforms as
\[
  \lambda_a\int\dd^D x\,O_a(x)
  \mapsto
  \lambda_a b^{D-\Delta_a}
  \int\dd^D x\,O_a(x).
\]
Comparing this scaling of the coupling with
\(\lambda_a(b)=b^{y_a}\lambda_a\) gives \(y_a=D-\Delta_a\).  For a Jordan
block, solving the linearized RG equation gives the same power \(b^{y_a}\)
multiplied by polynomials in \(\log b\), which is the RG form of logarithmic
operator mixing.
\end{proof}

\section{RG Monotones and Fixed-Point Ordering}
\label{sec:rg-monotones-fixed-point-ordering}

Let \(\mathfrak T\) be a class of unitary Lorentzian QFTs equipped with
renormalized stress-tensor source functionals and let an RG trajectory in
\(\mathfrak T\) have a UV endpoint \(\mathcal C_{\rm UV}\) and an IR endpoint
\(\mathcal C_{\rm IR}\), each a CFT in the sense of the preceding section.
The length scale \(R\) increases toward the infrared.  An endpoint monotone is
a real number \(\mathfrak M(\mathcal C)\), assigned to every endpoint CFT in
the class, together with an interpolating quantity \(\mathfrak M(R)\) on the
trajectory such that
\[
  \lim_{R\to0}\mathfrak M(R)=\mathfrak M(\mathcal C_{\rm UV}),
  \qquad
  \lim_{R\to\infty}\mathfrak M(R)=\mathfrak M(\mathcal C_{\rm IR}),
  \qquad
  \frac{\dd}{\dd\log R}\mathfrak M(R)\leq0 .
\]
The endpoint consequence is
\(\mathfrak M(\mathcal C_{\rm UV})\geq
\mathfrak M(\mathcal C_{\rm IR})\).  This definition keeps the mathematical
content in the class \(\mathfrak T\): positivity, stress-tensor existence,
analyticity, infrared behavior, and contact-term conventions are hypotheses
on the trajectory, not decorations on the conclusion.

In two spacetime dimensions the fixed-point number is the Virasoro central
charge \(c_{2d}\).  Under the hypotheses used in the
Zamolodchikov--Polchinski formulation--unitarity, Poincare invariance,
a normalizable vacuum, a local stress tensor, reflection positivity or
Hilbert-space positivity, and regularity of the stress-tensor two-point
distributions--write the Euclidean two-point functions in complex coordinates
as
\begin{align}
  \left\langle T(z,\bar z)T(0)\right\rangle
  &=
  \frac{F(R)}{z^4},
  \nonumber\\
  \left\langle T(z,\bar z)\Theta(0)\right\rangle
  &=
  \frac{G(R)}{z^3\bar z},
  \nonumber\\
  \left\langle \Theta(z,\bar z)\Theta(0)\right\rangle
  &=
  \frac{H(R)}{z^2\bar z^2},
  \qquad
  R=|z|,
  \qquad
  \Theta=T^\mu{}_\mu .
  \label{eq:zamolodchikov-form-factors}
\end{align}
The powers of \(z,\bar z\) isolate the tensor weights, while the functions
\(F,G,H\) depend only on \(R\).  Conservation of the stress tensor gives the
Zamolodchikov combination
\begin{equation}
  C(R)=2F(R)-G(R)-\frac38H(R)
  \label{eq:zamolodchikov-c-function}
\end{equation}
with derivative
\begin{equation}
  \frac{\dd C(R)}{\dd\log R}
  =
  -\frac34H(R)\leq0 .
  \label{eq:two-dimensional-c-function-derivative}
\end{equation}
Reflection positivity gives \(H(R)\geq0\).  At a conformal endpoint
\(\Theta=0\) at separated flat-space points, \(G=H=0\) and
\(F=c_{2d}/2\) in the normalization
\(\langle T(z)T(0)\rangle=c_{2d}/(2z^4)\), so \(C(R)\) tends to the central
charge of that endpoint.  Thus \(c_{\rm UV}\geq c_{\rm IR}\) for trajectories
satisfying the displayed hypotheses.

In three spacetime dimensions a fixed-point quantity of central importance is
the finite sphere free energy
\[
  F_{\mathcal C}
  =
  -\log |Z_{\mathcal C}(S^3)| ,
\]
after the local counterterms allowed on \(S^3\) have been fixed.  This
quantity is part of conformal data once a scheme for the curved-background
partition function has been declared.  Monotonicity statements involving
\(F_{\mathcal C}\) require additional input about the RG trajectory and its
regulated observables; they are not used here as definitions of a CFT.  Gauge
constraints, zero modes, topological sectors, and background counterterms must
be included in the definition of \(Z(S^3)\).

In four spacetime dimensions the endpoint quantity is the Euler coefficient
\(a_{\rm W}\) in the Weyl anomaly,
\[
  \left\langle T^\mu{}_\mu\right\rangle_g
  =
  \frac{1}{16\pi^2}
  \left(
    c_{\rm W}W^2-a_{\rm W}E_4+b_{\rm D}\nabla^2R
  \right),
\]
with the notation developed in
Chapter~\ref{ch:volume-iii-stress-tensor-weyl}.  The standard
Komargodski--Schwimmer construction should be read here as a conditional
dilaton-dispersion argument, not as a theorem-level proof from the general
axioms of four-dimensional QFT.  Its input is a unitary Lorentz-invariant
flow between CFT endpoints, a renormalized stress-tensor source functional
along the flow, a weak spectator dilaton \(\tau\) coupled by
\(g_{\mu\nu}=\ee^{-2\tau}\eta_{\mu\nu}\), a controlled large-\(f\) decoupling
limit for the spectator sector, and an infrared prescription under which the
dilaton scattering amplitude exists.  Under these hypotheses,
Wess--Zumino consistency for the mismatch of the UV and IR Weyl anomalies
fixes the flat-space four-derivative dilaton interaction whose on-shell
four-dilaton amplitude is
\begin{equation}
  \mathcal A(s,t,u)
  =
  \frac{a_{\rm UV}-a_{\rm IR}}{f^4}
  \left(s^2+t^2+u^2\right)
  +O(p^6),
  \qquad
  s+t+u=0 .
  \label{eq:ks-four-dilaton-amplitude}
\end{equation}
In the forward limit \(t=0\), this becomes
\[
  \mathcal A(s,0,-s)
  =
  \frac{2(a_{\rm UV}-a_{\rm IR})}{f^4}s^2+O(s^4).
\]
If, in addition, the forward amplitude obeys the analyticity, crossing,
polynomial boundedness, unitarity, and infrared-regulator conditions needed
for the dispersion relation, and if the large-circle and small-circle
contour contributions are controlled, Cauchy's theorem applied to
\(\mathcal A(s)/s^3\) gives the formal sum rule
\begin{equation}
  a_{\rm UV}-a_{\rm IR}
  =
  \frac{f^4}{\pi}
  \int_0^\infty\dd s\,
  \frac{\sigma_{\tau\tau}(s)}{s^2},
  \label{eq:ks-a-theorem-sum-rule}
\end{equation}
where the optical theorem has been used in the form
\(\operatorname{Im}\mathcal A(s,0)=s\,\sigma_{\tau\tau}(s)\).  Under the same
scattering assumptions, unitarity gives \(\sigma_{\tau\tau}(s)\geq0\), hence
\[
  a_{\rm UV}\geq a_{\rm IR}.
\]
If such a positive cross-section representation is available with a scale
cutoff, the same formal sum rule defines an interpolating quantity
\begin{equation}
  a(\mu)
  =
  a_{\rm UV}
  -
  \frac{f^4}{\pi}
  \int_\mu^\infty\dd s\,
  \frac{\sigma_{\tau\tau}(s)}{s^2},
  \qquad
  \frac{\dd a(\mu)}{\dd\mu}
  =
  \frac{f^4}{\pi}
  \frac{\sigma_{\tau\tau}(\mu)}{\mu^2}
  \geq0 .
  \label{eq:ks-interpolating-a}
\end{equation}
Thus \(a(\mu)\) increases with the energy scale \(\mu\) and decreases with
the length scale \(R\sim\mu^{-1}\), inside the same conditional scattering
framework.  This \(a(\mu)\) is not by itself a nonperturbative local RG
\(c\)-function constructed for every four-dimensional QFT.  The coefficient
\(c_{\rm W}\) does not obey an analogous general endpoint inequality; its role
in fixed-point data is through stress-tensor correlators, not through this
dilaton sum rule.

\begin{remark}[Status of the monotonicity statements]
The preceding paragraphs state the monotonicity statements at the level at
which this volume uses them.  The two-dimensional result is a theorem after
the analytic assumptions entering the \(C(R)\) construction have been fixed.
The three-dimensional sphere quantity is recorded here as fixed-point data;
the core CFT volume does not use an entropic construction as a structural
input.  The four-dimensional dilaton discussion is an inspired and highly
constraining physical argument whose conclusion follows only after the
displayed source-functional, decoupling, infrared, analyticity, crossing,
boundedness, and positivity assumptions have been supplied.  The monograph
does not treat it as a rigorous proof of a universal \(a\)-theorem, nor as a
substitute for constructing the RG trajectory or proving that its endpoints
satisfy the CFT hypotheses.\footnote{For the two-dimensional construction see
A.~B.~Zamolodchikov, ``Irreversibility of the flux of the renormalization
group in a 2D field theory,'' \emph{JETP Lett.} \textbf{43} (1986) 730--732,
and J.~Polchinski, ``Scale and conformal invariance in quantum field
theory,'' \emph{Nucl. Phys. B} \textbf{303} (1988) 226--236.  For the
four-dimensional dilaton argument see Z.~Komargodski and A.~Schwimmer,
``On renormalization group flows in four dimensions,'' \emph{JHEP}
\textbf{12} (2011) 099, arXiv:1107.3987.  For the three-dimensional sphere
partition-function quantity see D.~L.~Jafferis, I.~R.~Klebanov,
S.~S.~Pufu, and B.~R.~Safdi, ``Towards the \(F\)-theorem,'' arXiv:1103.1181.}
\end{remark}

\section{Conformal Field Theory as Local Fixed-Point Data}

The local data of a conformal field theory in flat \(D\)-dimensional space are
organized as follows.
\begin{enumerate}
  \item An algebraic vector space \(\mathcal V_{\rm loc}\) of renormalized
  local operators, graded by scaling dimension after diagonalizing the
  dilatation action.  Its elements are denoted
  \(\mathcal O_A(x)\), with \(A\) labeling a basis or a collection of
  multiplets.  In a unitary CFT this algebraic space maps, by radial
  quantization, to a dense subspace of the Hilbert space on \(S^{D-1}\); the
  Hilbert completion contains general sphere states and is not itself the
  space of local operators.
  \item A stress tensor \(T_{\mu\nu}\) satisfying conservation, symmetry, and
  the flat-space tracelessness condition
  \[
    \partial_\mu T^{\mu\nu}=0,
    \qquad
    T^{\mu\nu}=T^{\nu\mu},
    \qquad
    T^\mu{}_\mu=0 .
  \]
  \item An action of the conformal algebra on \(\mathcal V_{\rm loc}\),
  generated by the stress-tensor charges associated with conformal Killing
  fields.
  \item Euclidean separated-point correlation functions
  \[
    \left\langle
    \mathcal O_{A_1}(x_1)\cdots\mathcal O_{A_n}(x_n)
    \right\rangle,
    \qquad
    x_i\in\R^D,
  \]
  interpreted as distributions on the configuration space
  \(x_i\neq x_j\), together with contact-term extensions specified by the
  source Ward identities.
\end{enumerate}
For unitary Lorentzian theories these Euclidean distributions are related to
Wightman distributions by analytic continuation and reflection positivity,
under the same reconstruction hypotheses used for ordinary Euclidean Green
functions.

\section{Source Functionals and Ward-Identity Contact Terms}
\label{sec:cft-source-functionals-ward-contacts}

The separated-point correlators in the preceding list are restrictions of
distributions obtained from a source functional.  Work in Euclidean signature
for this construction.  The fixed-point functional used here is the
renormalized metric-and-operator source limit constructed in
Section~\ref{sec:background-metric-source-functional}, specialized to a point
of the RG flow where the beta functions vanish and to a stress-tensor
improvement convention appropriate to the fixed point.  Thus \(g_{\mu\nu}\) is
a smooth background metric, \(J^A(x)\) are smooth compactly supported
sources for a chosen local operator basis \(\mathcal O_A\), and local
functionals of \(g\) and \(J\) specify the contact-term chart.  The sign
normalization of \(W_\ast\) is the Euclidean free-energy, or effective-action,
normalization \(W_\ast=-\log Z_\ast\).  The connected-log functional is denoted
\(K_\ast=\log Z_\ast\) when it is used.

\begin{definition}[Status of fixed-point source brackets]
\label{def:cft-fixed-point-source-bracket-status}
The fixed-point partition functional
\[
  Z_\ast[g,J]:=\exp\{-W_\ast[g,J]\}
\]
is source-functional data of the fixed point.  The symbol
\(\langle\cdots\rangle_{g,\ast}\) used below denotes evaluation by this
curved-background source functional.  A literal Borel measure or an
unregulated continuum path integral would require an additional construction.
If a presentation of the CFT supplies only flat separated-point correlators,
then a compatible curved-background functional
\(Z_\ast[g,J]\), including its contact-term and Weyl-anomaly extension, is an
additional datum or theorem to be stated before the metric derivatives below
are used.  The bracket notation
\[
  \left\langle
  \exp\left(\int\dd^Dx\,\sqrt g\,J^A(x)\mathcal O_A(x)\right)
  \right\rangle_{g,\ast}
  :=
  Z_\ast[g,J]
\]
is a shorthand for that data.  Its mathematical construction depends on the
framework in which the fixed point has been supplied:
\begin{enumerate}[label=(\roman*)]
  \item in a constructive Euclidean model, \(Z_\ast[g,J]\) is obtained from
  the constructed measure or generalized random field after the local
  source counterterms have been fixed;
  \item in a lattice or Wilsonian construction, \(Z_\ast[g,J]\) is the
  continuum limit of regulated source functionals in a specified topology of
  distributions, after tuning and renormalizing the source chart;
  \item in an abstract CFT presentation, \(Z_\ast[g,J]\) is a generating
  functional for a compatible family of Euclidean Schwinger distributions and
  contact-term extensions satisfying the stated Ward identities, positivity
  or reflection-positivity hypotheses when unitarity is required, and the
  reconstruction assumptions used in this volume;
  \item in perturbative fixed-point calculations, \(Z_\ast[g,J]\) is a formal
  power series whose coefficients are renormalized source distributions.
\end{enumerate}
Thus the brackets in this section are typed by the chosen fixed-point
construction.  They are expectation brackets when a measure construction has
been provided, continuum-limit brackets when a regulated limit has been
proved or assumed as part of the data, and distributional generating
brackets in an axiomatic or perturbative presentation.
\end{definition}

\begin{hypothesis}[Existence of the fixed-point source chart]
\label{hyp:cft-fixed-point-source-chart-existence}
All differentiations of \(W_\ast[g,J]\) below are made in a source chart with
the following properties.  There is an open neighborhood \(\mathcal U_\ast\)
of \((g,J)=(\delta,0)\) in the Frechet space of smooth Euclidean metrics equal
to \(\delta_{\mu\nu}\) outside a compact set and smooth compactly supported
operator sources.  The functional \(W_\ast\) is a real-valued functional on
\(\mathcal U_\ast\), defined modulo local functionals of \(g\) and \(J\), whose
Gateaux derivatives of all orders exist as distributions.  More explicitly,
for every \((g,J)\in\mathcal U_\ast\) and every finite list of compactly
supported metric and source variations
\[
  u_r=(h^{(r)}_{\mu\nu},j^{(r)A}),\qquad r=1,\ldots,n,
\]
the multilinear functional
\[
  D_{u_1}\cdots D_{u_n} W_\ast[g,J]
\]
is continuous in the test-function topology and is represented by a
distribution on the corresponding Cartesian power of spacetime.

If the fixed point is supplied by a regulator, this hypothesis means that
there are regulated functionals \(Z_{\Lambda}[g,J]\) and local source
counterterms \(P_{\Lambda}[g,J]\) such that
\[
  W_{\Lambda}^{\rm ren}[g,J]
  =
  -\log Z_{\Lambda}[g,J]+P_{\Lambda}[g,J]
\]
has distributional limits of all such source derivatives on
\(\mathcal U_\ast\):
\[
  D_{u_1}\cdots D_{u_n}W_{\Lambda}^{\rm ren}[g,J]
  \longrightarrow
  D_{u_1}\cdots D_{u_n}W_\ast[g,J]
  \quad\hbox{in }\mathcal D' .
\]
Two regulators or subtraction prescriptions define the same source chart only
after a transition function has been specified; their difference is allowed to
be a local functional of the background fields, which changes only contact
terms in source derivatives.  Diffeomorphism and Weyl Ward identities for
\(W_\ast\) are additional properties of this chart, including the anomaly
functional when one is present.  They are not consequences of flat separated
correlators alone.
\end{hypothesis}

The definitions and Ward identities in the rest of this section are conditional
on Hypothesis~\ref{hyp:cft-fixed-point-source-chart-existence}.  In
perturbative constructions the hypothesis is a statement about formal power
series whose coefficients are renormalized source distributions.  In a
constructive, lattice, or Wilsonian construction it is a nonperturbative
source-dependent scaling-limit theorem when proved; otherwise it is part of
the CFT data being assumed.

For a regulated path-integral representative this convention reads
\[
  \exp\{-W_\ast[g,J]\}
  =
  \int \mathcal D_g\Phi\,
  \exp\left\{
    -S_\ast[\Phi;g]
    +\int\dd^D x\,\sqrt g\,J^A(x)\mathcal O_A(x)
  \right\},
\]
up to the local counterterms that define the source chart.  The minus sign in
the operator derivative below is forced by combining \(W_\ast=-\log Z_\ast\)
with the source term \(+\int J^A\mathcal O_A\) in the exponent; a source written
with the opposite sign would reverse this derivative convention.  In an abstract
CFT the displayed integral is replaced by the corresponding renormalized source
functional; the first derivatives below are the definition of the associated
contact-term convention.  This paragraph uses the Euclidean plus-source
coordinate.  Relative to the positive-measure convention
\(Z_E[J_E]=\int\exp\{-S_E-\langle J_E,\mathcal O\rangle\}\) of
Chapter~\ref{ch:volume-ii-generating-functionals-1pi}, the local operator
source here is \(J^A=-J_E^A\).  Relative to the Lorentzian convention
\(\exp\{\ii\langle J_L,\mathcal O\rangle_L\}\) and the lower-half-plane Wick
rotation of Chapter~\ref{ch:lorentzian-green-functions}, it is
\(J^A=J_L^{{\rm W},A}\), with the tensor/spin index Wick factors inserted when
the operator is not a scalar.  The source-coordinate sign is therefore fixed
before any Ward identity is differentiated.
The stress tensor and local operators are the source derivatives
\begin{equation}
  \left\langle T^{\mu\nu}(x)\right\rangle_{g,J}
  =
  -\frac{2}{\sqrt g(x)}
  \frac{\delta W_\ast[g,J]}{\delta g_{\mu\nu}(x)},
  \qquad
  \left\langle \mathcal O_A(x)\right\rangle_{g,J}
  =
  -\frac{1}{\sqrt g(x)}
  \frac{\delta W_\ast[g,J]}{\delta J^A(x)} .
  \label{eq:cft-source-derivative-conventions}
\end{equation}
The stress-tensor derivative is written with respect to \(g_{\mu\nu}\).  The
same convention may be written using \(g^{\mu\nu}\) by the chain-rule identity
\eqref{eq:stress-tensor-covariant-contravariant-derivatives}.
The sign in the first formula is the one induced by the Euclidean effective
action convention.  With
\[
  \delta_g S_\ast[\Phi;g]
  =
  -\frac12
  \int\dd^D x\,\sqrt g\,
  T^{\mu\nu}(x)\delta g_{\mu\nu}(x),
\]
one has \(\delta_g W_\ast=\langle\delta_g S_\ast\rangle\), hence
\[
  \delta_g W_\ast
  =
  -\frac12
  \int\dd^D x\,\sqrt g\,
  \left\langle T^{\mu\nu}(x)\right\rangle_{g,J}\delta g_{\mu\nu}(x).
\]
The operator-source derivative has the same overall minus sign for a different
reason: \(J^A\mathcal O_A\) appears with a plus sign in the exponent of
\(Z_\ast\), while \(W_\ast=-\log Z_\ast\).
Higher derivatives define Euclidean Schwinger distributions.  The source
functional is defined up to local functionals of the background sources.  This
is the continuum CFT version of the contact-term coordinate freedom described in
Section~\ref{sec:multiple-insertions-contact-coordinates}: the restriction to
the configuration space \(x_i\ne x_j\) is fixed by the linear operator
normalization, while the extension across collision diagonals is fixed by the
chosen local source chart.  The trace identity used in this chart is the
contact-distribution identity of
Proposition~\ref{prop:trace-identity-contact-distribution}; the Weyl-anomaly
functional and its Wess--Zumino consistency condition are part of the same
source data, not consequences of separated-point correlators alone.

The Ward identities are identities of these source-dependent distributions.
For an infinitesimal diffeomorphism generated by a compactly supported vector
field \(\xi^\mu\),
\[
  \delta_\xi g_{\mu\nu}
  =
  \nabla_\mu\xi_\nu+\nabla_\nu\xi_\mu,
  \qquad
  \delta_\xi J^A
  =
  \xi^\mu\nabla_\mu J^A
\]
for scalar sources.  Invariance of \(W_\ast[g,J]\) gives
\begin{equation}
  \nabla_\mu
  \left\langle T^{\mu\nu}(x)\right\rangle_{g,J}
  =
  \left\langle \mathcal O_A(x)\right\rangle_{g,J}
  \nabla^\nu J^A(x),
  \label{eq:cft-source-diffeomorphism-ward}
\end{equation}
with extra representation terms when the source carries spin indices.
Differentiating this identity with respect to \(J\) and then setting
\(J=0\) produces the contact terms in stress-tensor Ward identities.
Equivalently, for scalar insertions
\[
  X=\mathcal O_{A_1}(x_1)\cdots\mathcal O_{A_n}(x_n),
\]
the distributional translation Ward identity can be written in smeared form as
\begin{equation}
  \int\dd^D x\,
  \partial_\mu\xi_\nu(x)\,
  \left\langle T^{\mu\nu}(x)X\right\rangle
  =
  \sum_{i=1}^n
  \left\langle
    \mathcal O_{A_1}(x_1)\cdots
    (\xi^\rho\partial_\rho\mathcal O_{A_i})(x_i)\cdots
    \mathcal O_{A_n}(x_n)
  \right\rangle .
  \label{eq:cft-smeared-translation-ward}
\end{equation}
This equation is a compact way of recording the delta-function terms
supported at \(x=x_i\).

If the fixed point has no flat-space Weyl anomaly, the source chart for scalar
primary operators of dimensions \(\Delta_A\) may be chosen so that an
infinitesimal Weyl transformation satisfies
\[
  \delta_\sigma g_{\mu\nu}=2\sigma g_{\mu\nu},
  \qquad
  \delta_\sigma J^A=(\Delta_A-D)\sigma J^A .
\]
Differentiating the corresponding source Ward identity gives, for scalar
primaries and in flat space,
\begin{equation}
  \int\dd^D x\,\sigma(x)\,
  \left\langle T^\mu{}_\mu(x)X\right\rangle
  =
  -\sum_{i=1}^n
  \Delta_{A_i}\sigma(x_i)
  \left\langle X\right\rangle
  +
  \mathcal A_\sigma[X].
  \label{eq:cft-smeared-weyl-ward}
\end{equation}
The term \(\mathcal A_\sigma[X]\) is a local functional of the sources and
background metric.  It vanishes for flat separated-point correlators in a
Weyl-invariant source scheme, but it must be specified when correlators are
extended to coincident insertions or when the theory is placed on a curved
background.

A scalar primary operator \(\mathcal O\) has a scaling dimension
\(\Delta_{\mathcal O}\).  For scalar primaries
\(\mathcal O_i\), separated-point scale covariance is
\begin{equation}
  \left\langle
  \mathcal O_1(\lambda x_1)\cdots\mathcal O_n(\lambda x_n)
  \right\rangle
  =
  \lambda^{-\sum_i\Delta_i}
  \left\langle
  \mathcal O_1(x_1)\cdots\mathcal O_n(x_n)
  \right\rangle,
  \qquad
  \lambda>0 .
  \label{eq:cft-opening-scale-covariance}
\end{equation}
Operators with spin carry the corresponding finite-dimensional rotation or
Lorentz matrices.  The subsequent chapters derive these transformation laws
from stress-tensor charges.

\section{Ising Universality as a Fixed-Point Example}

The statistical Ising model is a Euclidean lattice model whose microscopic
state sum has a finite-dimensional diagonal trace representation.  Let
\(\Lambda\subset\mathbb Z^D\) be a finite lattice.  At every site
\(x\in\Lambda\) there is a spin variable \(s_x=\pm1\), represented in diagonal
operator notation by
\[
  V_x=\operatorname{span}\{\ket{\uparrow}_x,\ket{\downarrow}_x\},
  \qquad
  \widetilde{\Hilb}_\Lambda=\bigotimes_{x\in\Lambda}V_x,
\]
and by commuting diagonal operators \(\widehat S_x\) with eigenvalues
\(\pm1\).  With nearest-neighbor coupling normalized to one,
\[
  \widetilde H
  =
  -\sum_{\langle x x'\rangle}\widehat S_x\widehat S_{x'} .
\]
The thermal expectation of an observable \(\widehat O\) is
\begin{equation}
  \langle\widehat O\rangle_{\beta_{\rm T}}
  =
  \frac1Z
  \operatorname{Tr}_{\widetilde{\Hilb}_\Lambda}
  \left(
    \ee^{-\beta_{\rm T}\widetilde H}\widehat O
  \right),
  \qquad
  Z=
  \operatorname{Tr}_{\widetilde{\Hilb}_\Lambda}
  \ee^{-\beta_{\rm T}\widetilde H},
  \qquad
  \beta_{\rm T}=\frac1T .
  \label{eq:cft-opening-ising-thermal-trace}
\end{equation}
The tilde distinguishes the finite-dimensional trace space of the statistical
ensemble from the Hilbert space reconstructed from a continuum scaling limit.

Figure~\ref{fig:cft-opening-ising-source-figures} keeps track of the finite
Ising source construction used here: lattice spins, source insertions, and the
thermal trace belong to the regulated statistical ensemble before the continuum
CFT limit is taken.

\begin{figure}[H]
  \centering
  \resizebox{0.98\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small] at (1.35,3.0) {lattice spins};
      \foreach \i in {0,...,4} {
        \draw[gray!55] (0,\i*0.45)--(2.7,\i*0.45);
        \draw[gray!55] (\i*0.675,0)--(\i*0.675,1.8);
      }
      \foreach \x/\y/\s in {
        0/0/1,.675/0/-1,1.35/0/1,2.025/0/1,2.7/0/-1,
        0/.45/-1,.675/.45/1,1.35/.45/1,2.025/.45/-1,2.7/.45/1,
        0/.9/1,.675/.9/1,1.35/.9/-1,2.025/.9/1,2.7/.9/-1,
        0/1.35/-1,.675/1.35/1,1.35/1.35/1,2.025/1.35/1,2.7/1.35/-1,
        0/1.8/1,.675/1.8/-1,1.35/1.8/1,2.025/1.8/-1,2.7/1.8/1} {
        \ifnum\s=1
          \draw[green!50!black,->,thick] (\x,\y-0.15)--(\x,\y+0.15);
        \else
          \draw[green!50!black,->,thick] (\x,\y+0.15)--(\x,\y-0.15);
        \fi
      }
      \node[align=center] at (1.35,-0.45)
        {\(x=(i,j)\) in \(D=2\)\\\(s_x=\pm1\)};
    \end{scope}

    \begin{scope}[shift={(4.45,0)}]
      \node[font=\small] at (1.75,3.0) {magnetization};
      \draw[-{Latex[length=2mm]}, thick] (0,0)--(3.55,0) node[below] {\(\beta_{\rm T}\)};
      \draw[-{Latex[length=2mm]}, thick] (0,0)--(0,2.3) node[left] {\(|\langle s\rangle|\)};
      \draw[dashed,gray!70] (1.23,0)--(1.23,2.05);
      \node[below] at (1.23,0) {\(\beta_c\)};
      \draw[blue!75!black,very thick]
        (0.05,0.02)--(1.23,0.02)
        .. controls (1.30,0.38) and (1.65,1.30) .. (2.45,1.78)
        .. controls (2.85,2.00) and (3.2,2.05) .. (3.45,2.05);
      \node[blue!75!black,align=center] at (2.65,1.15)
        {ordered\\phase};
      \node[gray!70!black] at (-0.18,2.05) {\(1\)};
    \end{scope}

    \begin{scope}[shift={(9.45,0)}]
      \node[font=\small] at (1.75,3.0) {spin two-point function};
      \draw[-{Latex[length=2mm]}, thick] (0,0)--(3.8,0) node[below] {\(r\)};
      \draw[-{Latex[length=2mm]}, thick] (0,0)--(0,2.3) node[left] {\(f(r)\)};
      \draw[green!50!black, very thick]
        (0.15,2.02)..controls (0.85,1.72) and (1.75,1.48)..(3.55,1.46);
      \draw[blue!75!black, very thick]
        (0.15,2.02)..controls (0.55,1.28) and (1.25,0.70)..(3.35,0.28);
      \draw[purple!75!black, very thick]
        (0.15,2.02)..controls (0.45,0.95) and (0.95,0.28)..(2.25,0.05);
      \node[green!50!black] at (2.85,1.78) {\(T<T_c\)};
      \node[blue!75!black] at (2.65,0.73) {\(T=T_c\)};
      \node[purple!75!black] at (1.45,0.43) {\(T>T_c\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The Ising fixed point is reached by tuning the temperature.  The
  finite lattice trace space defines the statistical ensemble, spontaneous
  magnetization appears for \(T<T_c\), and the critical two-point function has
  power-law decay.}
  \label{fig:cft-opening-ising-source-figures}
\end{figure}

In \(D=2\), write \(S_{N,M}\) for the spin observable at
\(x=(N,M)\) and set
\[
  f_{N,M}
  =
  \langle S_{0,0}S_{N,M}\rangle_{\beta_{\rm T}}
  \simeq f(r),
  \qquad
  r=(N^2+M^2)^{1/2},
\]
for \(r\gg1\) and \(T\) close to \(T_c\).  The long-distance regimes are
\[
  f(r)\longrightarrow |\langle s\rangle|^2>0
  \qquad (T<T_c),
\]
\[
  f(r)\sim \ee^{-r/\xi(T)}
  \qquad (T>T_c),
  \qquad
  \xi(T)\sim (T-T_c)^{-\nu},
\]
and
\begin{equation}
  f(r)\sim r^{-2\Delta_\sigma}
  \qquad (T=T_c).
  \label{eq:cft-opening-ising-critical-power}
\end{equation}
The first numerical values to keep in view are
\[
\begin{array}{c|cc}
  D & \Delta_\sigma & \nu \\ \hline
  2 & \frac18 & 1\\[2pt]
  3 & 0.5181488\ldots & 0.629971\ldots
\end{array}
\]
where \(\Delta_\sigma\) is the scaling dimension of the continuum spin field
and \(\nu\) is the correlation-length exponent.
\begin{remark}[Numerical provenance in three dimensions]
The \(D=3\) entries come from numerical conformal-bootstrap computations.  They
are obtained by applying crossing symmetry and reflection-positivity constraints
to mixed four-point functions of
\(\sigma\), \(\varepsilon\), and \(T_{\mu\nu}\), followed by
finite-dimensional semidefinite relaxations of the resulting bootstrap
inequalities and numerical stability tests as the relaxation order changes.
In such a computation the assumptions are the existence of a
unitary three-dimensional CFT with the \(\mathbb Z_2\) quantum numbers and
operator gaps imposed in the search; the output is a small allowed region for
low-lying CFT data.  A constructive proof of the lattice scaling limit is a
separate theorem.
Chapter~\ref{ch:volume-iii-operator-product-expansion} develops the OPE,
reflection-positivity, and crossing equations used by this method, but the
present chapter records only the numerical CFT data and their provenance.  A
recent mixed-correlator computation including the stress tensor gives
\(\Delta_\sigma=0.518148806(24)\) and
\(\Delta_\varepsilon=1.41262528(29)\) for the leading
\(\mathbb Z_2\)-odd and \(\mathbb Z_2\)-even scalar primaries of the critical
Ising CFT; using \(\Delta_\varepsilon=D-1/\nu\) at \(D=3\) gives
\(\nu=0.629970975\ldots\).  The source is Chang, Dommes, Erramilli, Homrich,
Kravchuk, Liu, Mitchell, Poland, and Simmons-Duffin,
\emph{Bootstrapping the 3d Ising Stress Tensor}, JHEP \textbf{03} (2025) 136,
arXiv:2411.15300.
\end{remark}

\begin{definition}[Universal scaling functions in the Ising setting]
\label{def:cft-opening-universal-scaling-functions}
Fix a class \(\mathfrak R\) of reflection-positive microscopic regulators with
\(\mathbb Z_2\) symmetry, one spinlike relevant direction, and one
temperaturelike relevant direction.  For each regulator \(R\in\mathfrak R\),
let \(a_R\) be the lattice spacing, let \(p_R\) denote microscopic couplings
near the critical surface, and let \(X_A^{(R)}(x)\) be lattice observables
whose leading scaling fields are \(O_A\).  A separated \(n\)-point scaling
function is a distributional limit on the configuration space \(x_i\neq x_j\)
of the form
\[
  \mathcal F_{A_1\ldots A_n}(x_1,\ldots,x_n;s)
  =
  \lim_{a_R\to0}
  \prod_{i=1}^n Z_{A_i}^{(R)}
  \left\langle
    X_{A_1}^{(R)}(a_R^{-1}x_1)\cdots
    X_{A_n}^{(R)}(a_R^{-1}x_n)
  \right\rangle^{\rm conn}_{p_R(s)} ,
\]
where \(s=(s_t,s_h)\) are fixed continuum scaling coordinates for the thermal
and magnetic relevant deformations, and \(Z_A^{(R)}\) are multiplicative
normalizations of the scaling fields.  The scaling function is called
universal for \(\mathfrak R\) when any two regulators in \(\mathfrak R\) give
the same \(\mathcal F\) after analytic changes of the relevant coordinates
\((p_R-p_{R,c})\mapsto(s_t,s_h)\), multiplicative field normalizations, and
subtraction of analytic background terms in the free energy.  This definition
concerns separated-point kernels and their finite-size or thermodynamic
limits.  Critical temperatures, lattice metric factors, operator
normalizations, analytic free-energy backgrounds, and contact-term extensions
are regulator data rather than universal scaling functions.
\end{definition}

\begin{proposition}[Conditional determination by complete CFT data]
\label{prop:cft-opening-ising-scaling-functions-conditional}
Assume that a regulator class \(\mathfrak R\) has a continuum critical limit
which is a unitary Euclidean CFT \(\mathcal C_{\rm Ising}\), that the lattice
spin and energy observables converge to the primary fields \(\sigma\) and
\(\varepsilon\) after the normalizations in
Definition~\ref{def:cft-opening-universal-scaling-functions}, and that the
OPE of \(\mathcal C_{\rm Ising}\) converges as a distribution on every
separated Euclidean configuration.  Then the critical separated-point scaling
functions are determined by the complete CFT data of \(\mathcal C_{\rm Ising}\):
the primary spectrum with spin and \(\mathbb Z_2\) parity, two-point
normalizations, and OPE coefficients.

If, in addition, the relevant deformation
\[
  S_{\mathcal C}
  \longmapsto
  S_{\mathcal C}
  +
  s_t\int\dd^Dx\,\varepsilon(x)
  +
  s_h\int\dd^Dx\,\sigma(x)
\]
is constructed as a continuum source-dependent QFT with a specified
renormalized contact-term chart, then the off-critical universal scaling
functions in the variables \(s_t,s_h\) are the separated connected correlators
of this deformed continuum theory.  Their short-distance expansion near the
critical point is obtained by conformal perturbation theory from the complete
CFT data and the chosen deformation scheme.  Their global massive scaling
functions require the deformed continuum theory itself, because finitely many
critical exponents such as \(\Delta_\sigma\) and \(\nu\) do not determine all
massive correlation functions.
\end{proposition}

\begin{proof}
The convergence hypothesis identifies the critical scaling limit of the
normalized lattice observables with the separated-point correlators of
\(\mathcal C_{\rm Ising}\).  The complete CFT data determine those correlators
as follows.  In a ball of configuration space where \(x_1\) approaches \(x_2\)
with all other insertions separated, OPE convergence gives
\[
  O_{A_1}(x_1)O_{A_2}(x_2)
  =
  \sum_B
  C_{A_1A_2}{}^B(x_{12},\partial_{x_2})\,O_B(x_2)
\]
as an equality of distributions after pairing with test functions supported in
that ball.  The coefficient distributions
\(C_{A_1A_2}{}^B\) are fixed by the dimensions, spins, representation labels,
two-point normalizations, and OPE coefficients.  Iterating the OPE reduces an
\(n\)-point function on such a configuration-space chart to a convergent sum of
two- and three-point tensor structures with the same data.  Compatibility of
the convergent OPE expansions on overlapping charts is part of the CFT data
through associativity/crossing.  Hence the critical separated-point kernels are
fixed by the complete CFT data.

For the deformed statement, the assumed continuum source-dependent QFT supplies
a generating functional \(W_{\mathcal C}[s_t,s_h;J]\).  Its functional
derivatives with respect to the sources \(J^A\), restricted to separated
points, are precisely the off-critical scaling functions in the chosen
normalization and scaling-coordinate chart.  Expanding
\(\exp[-s_t\int\varepsilon-s_h\int\sigma]\) near \(s_t=s_h=0\) expresses the
short-distance coefficients as integrated CFT correlators.  The specified
renormalized contact-term chart is the datum that defines those integrals at
coincident integration points.  The full finite-\(s\) functions are therefore
correlators of the relevantly deformed continuum theory.
\end{proof}

\begin{remark}[Status of the three-dimensional Ising scaling functions]
\label{rem:cft-opening-ising-scaling-function-status}
For \(D=2\), the conformal data are exact data of the Ising minimal model, and
many scaling limits of the planar lattice model are theorem-level objects.  For
\(D=3\), the statement that the nearest-neighbor Ising model and the scalar
\(\mathbb Z_2\)-invariant \(\phi^4\) regulator class converge to a single
unitary Ising CFT with the scaling functions of
Definition~\ref{def:cft-opening-universal-scaling-functions} is a
mathematically open continuum-limit statement.  The numerical values displayed
above are bootstrap estimates for the candidate CFT data.  The parenthetical
uncertainties attached to
\(\Delta_\sigma=0.518148806(24)\) and
\(\Delta_\varepsilon=1.41262528(29)\) are numerical uncertainties reported by
the finite semidefinite bootstrap computation; they are neither convergence
rates for the lattice spacing \(a_R\to0\) nor rigorous error bounds for a
constructive lattice-to-CFT theorem.  If a leading irrelevant exponent
\(\omega>0\) and a regulator-dependent coefficient are supplied by a separate
RG analysis, one may write corrections of the form \(O(a_R^\omega)\) for
specified observables, but such a rate is additional data beyond the two
numbers in the table.
\end{remark}

\begin{remark}[Existence status in \(D\geq4\)]
\label{rem:cft-opening-dgefour-existence}
The two-row table above records Ising fixed-point data in \(D=2,3\).  For
\(D\geq4\), the existence question separates into the following cases.  A free
conformal field theory means a local conformal theory generated by fields
obeying linear equations of motion whose vacuum correlators satisfy Wick
factorization; its local operator space can be large, but its separated
connected correlators are fixed by two-point data.  A dynamically interacting
CFT means a stress-tensor CFT whose separated correlators are not generated by
Wick factorization from free conformal sectors and decoupled topological
sectors.

There are explicit unitary Gaussian CFTs in \(D\geq4\): the improved massless
scalar and massless Dirac field in every allowed integer dimension, and the
Maxwell field in \(D=4\).  These theories supply exact local conformal data, but
they do not supply interacting fixed-point data.  The Wilson--Fisher
\(\epsilon\)-expansion near \(D=4\) supplies formal perturbative conformal data
at \(D=4-\epsilon\).  For nonintegral \(D\), the usual Hilbert-space positivity
condition of an integer-dimensional Lorentzian QFT is not literally available,
and the analytic continuation of conformal representation theory contains
negative-norm states in generic nonintegral dimension.  The
\(\epsilon\)-expansion is therefore a powerful asymptotic computation of local
CFT data, not by itself a construction of a unitary CFT at a fixed integer
dimension.

For the positive-measure scalar route, four dimensions are also constrained by
triviality theorems for critical Ising and \(\phi^4_4\)-type scaling limits:
within those hypotheses the continuum limits are Gaussian.  This result blocks
the simplest scalar construction of an interacting \(D=4\) CFT; it does not
classify all possible four-dimensional fixed points.  Gauge-theory mechanisms
such as Banks--Zaks fixed points and later supersymmetric conformal gauge
theories provide physically controlled fixed-point constructions under their
own assumptions.  In this core CFT volume, any such \(D\geq4\) interacting
example is used only after its stress tensor, unitarity/positivity, locality,
operator spectrum, and conformal Ward identities have been specified as part of
the data.
\footnote{For nonintegral-dimensional unitarity issues in the
Wilson--Fisher expansion, see M.~Hogervorst, S.~Rychkov, and B.~C.~van Rees,
``Unitarity violation at the Wilson-Fisher fixed point in
\(4-\epsilon\) dimensions,'' arXiv:1512.00013.  For four-dimensional scalar
triviality, see M.~Aizenman and H.~Duminil-Copin, ``Marginal triviality of the
scaling limits of critical 4D Ising and \(\phi^4_4\) models,''
\emph{Ann. of Math.} \textbf{194} (2021) 163--235.  For the perturbative
gauge-theory fixed-point mechanism, see T.~Banks and A.~Zaks, ``On the phase
structure of vector-like gauge theories with massless fermions,''
\emph{Nucl. Phys. B} \textbf{196} (1982) 189--204.}
\end{remark}

The same fixed-point basin can be represented by a scalar Euclidean lattice
field theory, in the sense of
Definition~\ref{def:cft-opening-universal-scaling-functions}.  Replace the
two-valued spin by a real variable \(\phi_x\), use a \(\mathbb Z_2\)-even
double-well single-site potential, and retain the finite Brillouin-zone
cutoff.  At long distance its action has the form
\begin{equation}
  S_{\rm lat}[\phi]
  =
  \int\dd^Dx\,
  \left[
    \frac12 Z_\phi(\partial_\mu\phi)^2
    +
    V(\phi)
    +
    \sum_A c_A\,\mathcal O_A^{\rm aniso}(\phi,\partial\phi)
  \right],
  \label{eq:cft-opening-ising-scalar-cutoff}
\end{equation}
where the last sum consists of higher-derivative operators allowed by the
hypercubic lattice symmetry.  These terms display the finite cutoff and the
breaking of \(SO(D)\) rotations at the microscopic scale.

At \(T=T_c\), equivalently at the tuned massless critical point, the
correlation length is infinite in lattice units.  The separated long-distance
correlators are described by the Ising conformal field theory.  For
\(D=2,3\), this is the infrared fixed point of the \(\mathbb Z_2\)-invariant
massless scalar \(\phi^4\) fixed-point basin.  The continuum fields obey the
dictionary
\[
  \phi \longmapsto \sigma,
  \qquad
  [\phi^2]\longmapsto \varepsilon ,
\]
where \(\sigma\) is the spin field and \(\varepsilon\) is the thermal
energy-density field after identity mixing has been subtracted.  The thermal
deformation
\[
  \Delta S_t=\int\dd^Dx\,u_t\,\varepsilon(x)
\]
has RG eigenvalue \(y_t=D-\Delta_\varepsilon\).  Since
\(\xi^{-1}\propto |u_t|^{1/y_t}\), comparison with
\(\xi\propto |T-T_c|^{-\nu}\) gives
\begin{equation}
  \Delta_\varepsilon
  =
  D-\frac1{\nu}.
  \label{eq:cft-opening-ising-energy-dimension}
\end{equation}

Figure~\ref{fig:cft-opening-ising-scalar-cutoff} records the scalar cutoff
presentation used to compare the Ising scaling direction with a local
continuum field coordinate.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small] at (0,2.65) {cutoff scalar model};
      \draw[-{Latex[length=2mm]}, thick] (-1.35,0)--(1.45,0) node[below] {\(\phi\)};
      \draw[-{Latex[length=2mm]}, thick] (0,-0.2)--(0,2.05) node[left] {\(V(\phi)\)};
      \draw[blue!70!black,very thick]
        plot[smooth] coordinates {(-1.25,1.85) (-0.72,0.35) (0,1.05) (0.72,0.35) (1.25,1.85)};
      \node[blue!70!black] at (0,-0.55) {\(\mathbb Z_2\)-even};
    \end{scope}

    \begin{scope}[shift={(4.1,0)}]
      \node[font=\small] at (0,2.65) {finite momentum cutoff};
      \draw[thick] (-1.15,-1.15) rectangle (1.15,1.15);
      \draw[-{Latex[length=2mm]}, gray!70] (-1.55,0)--(1.65,0) node[below] {\(k^1\)};
      \draw[-{Latex[length=2mm]}, gray!70] (0,-1.55)--(0,1.65) node[left] {\(k^2\)};
      \node at (-1.15,-1.42) {\(-\pi\)};
      \node at (1.15,-1.42) {\(\pi\)};
      \node[align=center] at (0,-1.9) {Brillouin zone};
    \end{scope}

    \draw[-{Latex[length=2mm]}, thick] (5.75,0.55)--(7.05,0.55);
    \node[align=center] at (6.40,1.05) {tune\\mass};

    \begin{scope}[shift={(8.05,0)}]
      \node[font=\small] at (0,2.65) {infrared fixed point};
      \draw[thick,blue!60!black] (-1.2,0.15)..controls (-0.55,1.7)
        and (0.55,1.7)..(1.2,0.15);
      \draw[thick,blue!60!black,dashed] (-1.2,0.15)..controls (-0.55,-0.38)
        and (0.55,-0.38)..(1.2,0.15);
      \fill[purple!75!black] (-0.42,0.75) circle (2pt);
      \node[purple!75!black,above] at (-0.42,0.86) {\(\sigma\)};
      \fill[orange!85!black] (0.55,0.58) circle (2pt);
      \node[orange!85!black,above] at (0.55,0.69) {\(\varepsilon\)};
      \node[align=center] at (0,-0.70)
        {\(\Delta_\sigma\), \(\Delta_\varepsilon=D-1/\nu\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The scalar cutoff representation of the Ising fixed-point basin
  has a finite Brillouin-zone cutoff and lattice anisotropy.  At the tuned
  critical point its long-distance operator data are those of the Ising
  conformal field theory.}
  \label{fig:cft-opening-ising-scalar-cutoff}
\end{figure}
