\chapter{Radial Quantization and the State-Operator Correspondence}
\label{ch:volume-iii-radial-quantization}
\ConventionsMixedCFT


\section{Local Correlators as the Primary Data}

Let \(\widehat{\mathcal O}_I(x)\) denote a Lorentzian local operator-valued
distribution on \(\R^{1,D-1}\).  The label \(I\) includes all internal and
spin indices.  Local commutativity means that for spacelike separated points
\[
  [\widehat{\mathcal O}_I(x),\widehat{\mathcal O}_J(y)]_\pm=0,
  \qquad
  (x-y)^2>0,
\]
where the graded commutator is used when fermionic operators are present.
For a vacuum vector \(\ket{\vac}\), the unordered Lorentzian correlation
functions are the Wightman distributions
\[
  W_{I_1\cdots I_n}(x_1,\ldots,x_n)
  =
  \bra{\vac}
  \widehat{\mathcal O}_{I_1}(x_1)\cdots
  \widehat{\mathcal O}_{I_n}(x_n)
  \ket{\vac}.
\]
Time-ordered Green functions are
\[
  G^T_{I_1\cdots I_n}(x_1,\ldots,x_n)
  =
  \bra{\vac}
  T\{\widehat{\mathcal O}_{I_1}(x_1)\cdots
  \widehat{\mathcal O}_{I_n}(x_n)\}
  \ket{\vac}.
\]
The corresponding Euclidean Schwinger functions are denoted by
\[
  S_{I_1\cdots I_n}(x^E_1,\ldots,x^E_n)
  =
  \left\langle
  \mathcal O_{I_1}(x^E_1)\cdots\mathcal O_{I_n}(x^E_n)
  \right\rangle_E .
\]
\begin{hypothesis}[Radial reconstruction data]
\label{hyp:radial-reconstruction-data}
The state--operator construction in this chapter is used for a Euclidean
Schwinger hierarchy \(S_{I_1\cdots I_n}\) satisfying the following conditions.
\begin{enumerate}
  \item The \(S_n\) are distributional correlation functions of renormalized
  local operators on \(\R^D\), with Euclidean covariance, graded permutation
  symmetry, and stated contact-term prescriptions at coincident points.
  \item Let \(\vartheta(x)=x/|x|^2\) be reflection through the unit sphere.  The
  operator basis carries an antilinear reflection map \(\Theta_{\rm rad}\) that
  acts on a smeared insertion supported in the unit ball by applying
  \(\vartheta\), the appropriate Weyl factor, and the spin representation of
  the reflected orthonormal frame.  For every finite linear combination \(F\)
  of such inside-supported insertions,
  \[
    \langle F,F\rangle_{\rm rad}
    =
    \left\langle \Theta_{\rm rad}F\,F\right\rangle_E
    \geq0 .
  \]
  The radial Hilbert space is the completion of this positive semidefinite
  form after quotienting by its null space.
  \item The one-parameter group generated by dilatations acts on this Hilbert
  space by a nonnegative self-adjoint operator \(D_{\rm rad}\).  Its spectrum
  on the finite-energy subspace is discrete, and for every \(E<\infty\) the
  spectral subspace with \(D_{\rm rad}\le E\) is finite-dimensional.  In
  particular every eigenspace has finite multiplicity, and the eigenvalue
  \(0\) subspace is spanned by the vacuum vector.  Thus the conformally
  invariant vacuum in the radial Hilbert space is unique.
  \item For each local operator \(\mathcal O_I\) used below, the radial limit of
  the insertion at the origin defines a finite-energy vector.  Descendant
  vectors are obtained by applying the translation generators, equivalently by
  derivatives of the insertion.  The local operator basis is radially complete:
  after quotienting by the null radial two-point pairing, origin insertions of
  local operators and their descendants span precisely the finite-energy
  subspace
  \[
    \mathcal H_{\rm fin}
    =
    \bigoplus_{\lambda<\infty}\ker(D_{\rm rad}-\lambda).
  \]
  Equivalently, every finite-energy eigenvector admits the local radial
  boundary-value data of
  Definition~\ref{def:local-radial-boundary-value-state} below, and hence has
  a representative as a local scaling field at the origin by
  Theorem~\ref{thm:operator-from-state-radial-boundary-value}, unique modulo
  null operators.
  \item The Euclidean correlators are related to Lorentzian Wightman
  distributions by the standard tube-domain analytic continuation, with the
  Lorentzian distributions satisfying the spectrum condition and local
  commutativity.  The flat-space vacuum correlators obey cluster decomposition:
  for two finite products of separated local insertions,
  \[
    A(x_1,\ldots,x_m)=\mathcal O_{I_1}(x_1)\cdots\mathcal O_{I_m}(x_m),
    \qquad
    B(y_1,\ldots,y_n)=\mathcal O_{J_1}(y_1)\cdots\mathcal O_{J_n}(y_n),
  \]
  and for a Euclidean vector \(a\) such that no coincident-point singularity is
  crossed along the translated configuration,
  \[
    \lim_{\lambda\to+\infty}
    \left\langle
      A(x_1+\lambda a,\ldots,x_m+\lambda a)\,
      B(y_1,\ldots,y_n)
    \right\rangle_E
    =
    \langle A(x_1,\ldots,x_m)\rangle_E\,
    \langle B(y_1,\ldots,y_n)\rangle_E .
  \]
\end{enumerate}
\end{hypothesis}

Under Hypothesis~\ref{hyp:radial-reconstruction-data}, radial reflection
positivity gives the inner product, \(D_{\rm rad}\) gives the cylinder
Hamiltonian, and local insertions at the origin give the complete
finite-energy state space.

The nonnegativity of \(D_{\rm rad}\) in
Hypothesis~\ref{hyp:radial-reconstruction-data} is the radial spectrum
condition for the vacuum representation.  It is not the statement, proved
later from descendant Gram positivity, that non-identity primary operators
obey the unitarity lower bounds of
Chapter~\ref{ch:volume-iii-unitarity-bounds-short-multiplets}.  The same
nonnegative self-adjoint operator \(D_{\rm rad}\) is used in the radial OPE
convergence theorem of
Chapter~\ref{ch:volume-iii-operator-product-expansion}.

These three versions are related by analytic continuation whenever the
Lorentzian theory satisfies the spectral and locality hypotheses needed for
the tube-domain construction.  For an ordered Wightman product, continue
\[
  x^0_k=-\ii x^D_{E,k}
\]
through a domain in which the imaginary parts remain ordered.  With the
convention \(x^0=-\ii\tau_E\), the ordering
\[
  \operatorname{Im}x^0_1<\operatorname{Im}x^0_2<\cdots<
  \operatorname{Im}x^0_n
\]
corresponds to
\[
  \tau_{E,1}>\tau_{E,2}>\cdots>\tau_{E,n}.
\]
Here \(\tau_E\) denotes Euclidean time inherited from Lorentzian analytic
continuation.  It is distinct from the cylinder coordinate \(\tau=\log r\)
introduced below.  The Euclidean correlator is the boundary value obtained
after this rotation.
Time-ordered Green functions arise from a contour whose real-time projection
places the insertions in chronological order.

\begin{definition}[Status of the complex-time contour functional]
\label{def:cft-contour-path-integral-status}
When a Lagrangian path-integral representative is available, the contour
notation used for this analytic continuation is tied to the Lorentzian
oscillatory convention of
Definition~\ref{def:lorentzian-oscillatory-path-integral} in
Chapter~\ref{ch:scalar-path-integrals-green-functions}.  Choose a finite
spatial regulator \(\Lambda\) and a finite subdivision \(P\) of the complex
time contour \(C\).  Let \(\mathcal C_{\Lambda,P}\) be the resulting
finite-dimensional space of retained bosonic variables, together with the
corresponding finite odd variables when fermions are present.  The contour
action is
\[
  S_{C,\Lambda,P}[\Phi]
  =
  \int_C \dd x^0\int\dd^{D-1}\mathbf x\,
  \mathcal L_{\Lambda,P}(\Phi,\partial\Phi),
\]
where the displayed integral is shorthand for the time-sliced finite
regulator action along \(C\).  For
\(F=\prod_i\mathcal O_i(x_i)\), the finite-regulator contour functional is
\begin{equation}
  I_{C,\Lambda,P}(F)
  =
  \int_{\mathcal C_{\Lambda,P}}
  d\nu_{\Lambda,P}(\Phi)\,
  \exp\{\ii S_{C,\Lambda,P}[\Phi]\}\,
  F_{\Lambda,P}(\Phi).
  \label{eq:regulated-complex-time-contour-functional}
\end{equation}
For even variables \(d\nu_{\Lambda,P}\) is an ordinary finite-dimensional
reference measure.  For odd variables it is a Berezin density.  For gauge
fields it is defined only after the finite-regulator gauge-fixing or BV data
and boundary conditions have been supplied.  If \(C\) contains Lorentzian
segments, \eqref{eq:regulated-complex-time-contour-functional} is an
oscillatory integral or oscillatory distribution, with Abel/Fourier boundary
value or stationary-phase meaning as in
Definition~\ref{def:lorentzian-oscillatory-path-integral}; Euclidean segments
contribute the corresponding heat-kernel damping in the same time-sliced
regulator.

The continuum symbol
\begin{equation}
  \left\langle\prod_i\mathcal O_i(x_i)\right\rangle_C
  :=
  \lim_{\Lambda,P}
  I_{C,\Lambda,P}
  \!\left(\prod_i\mathcal O_i(x_i)\right)
  \label{eq:complex-time-contour-continuum-symbol}
\end{equation}
means the specified distributional boundary value in a construction where the
limit exists.  In a perturbative construction it denotes the corresponding
formal asymptotic expansion obtained from the same regulator data.  In an
abstract CFT presentation the primary object is the Schwinger/Wightman
hierarchy and its tube-domain analytic continuation; the contour functional
is a representation of those boundary values.
\end{definition}

Figure~\ref{fig:cft-wightman-time-euclidean-contours} records the boundary-value
interpretation of complex-time contours used here: Euclidean, Lorentzian, and
radially ordered correlators are different boundary presentations of the same
analytic hierarchy once the relevant continuation exists.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      axis/.style={-{Latex[length=2mm]},thick},
      contour/.style={green!45!black,line width=0.95pt},
      point/.style={circle,fill=violet!75!black,inner sep=1.8pt}
    }
    \begin{scope}
      \node[font=\small] at (0,2.48) {Wightman contour};
      \draw[axis] (-2.45,0)--(2.55,0) node[right] {\(\operatorname{Re}x^0\)};
      \draw[axis] (0,-1.55)--(0,1.55)
        node[above left] {\(-\operatorname{Im}x^0\)};
      \draw[contour]
        (-2.10,-0.45) .. controls (-1.20,-0.30) and (-0.75,-0.95) ..
        (-0.25,-0.78) .. controls (0.20,-0.60) and (0.25,0.68) ..
        (0.80,0.72) .. controls (1.55,0.78) and (1.65,1.28) .. (1.80,1.50);
      \foreach \x/\y/\lab in {-1.55/-0.35/{x_3^0},-0.22/-0.78/{x_4^0},0.72/0.72/{x_1^0},1.48/0.95/{x_2^0}} {
        \node[point] at (\x,\y) {};
        \node[violet!75!black,below right] at (\x,\y) {\(\lab\)};
      }
    \end{scope}

    \begin{scope}[shift={(5.85,0)}]
      \node[font=\small] at (0,2.48) {time ordered};
      \draw[axis] (-2.20,0)--(2.30,0) node[right] {\(\operatorname{Re}x^0\)};
      \draw[axis] (0,-1.55)--(0,1.55)
        node[above left] {\(-\operatorname{Im}x^0\)};
      \draw[contour] (-2.0,-0.95)--(1.90,0.58);
      \foreach \x/\y/\lab in {-1.55/-0.77/{x_4^0},-0.78/-0.46/{x_3^0},0.18/-0.08/{x_2^0},1.05/0.27/{x_1^0}} {
        \node[point] at (\x,\y) {};
        \node[violet!75!black,above] at (\x,\y) {\(\lab\)};
      }
    \end{scope}

    \begin{scope}[shift={(11.35,0)}]
      \node[font=\small] at (0,2.48) {Euclidean};
      \draw[axis] (-1.65,0)--(1.65,0) node[right] {\(\operatorname{Re}x^0\)};
      \draw[axis] (0,-1.55)--(0,1.55)
        node[above left] {\(-\operatorname{Im}x^0=\tau_E\)};
      \draw[contour] (0,-1.25)--(0,1.45);
      \foreach \y/\lab in {1.05/{x_1^0},0.35/{x_2^0},-0.35/{x_3^0},-1.05/{x_4^0}} {
        \node[point] at (0,\y) {};
        \node[violet!75!black,right] at (0,\y) {\(\lab\)};
      }
    \end{scope}

    \draw[-{Latex[length=2mm]},orange!80!black,thick] (8.40,-1.70)
      .. controls (9.30,-2.18) and (10.30,-2.18) .. (11.20,-1.70);
    \node[orange!80!black] at (9.80,-2.25) {Wick rotation};
  \end{tikzpicture}
  }
  \caption{Different real-time orderings and Euclidean Schwinger functions are
  boundary values of the same analytic correlation functions, reached by
  deforming the complex-time contour without crossing singularities.}
  \label{fig:cft-wightman-time-euclidean-contours}
\end{figure}

\begin{remark}[Observables at a conformal fixed point]
At a conformal fixed point, the data used in this volume are local operators,
their correlation functions, and the Hilbert-space structure obtained from
radial quantization.  A separate scattering theory requires asymptotic
particle states.  Many conformal theories have no mass gap and no isolated
massive particle sectors, so their intrinsic observable data are organized by
local correlations rather than an \(S\)-matrix.  Four-dimensional Yang--Mills
theory illustrates a different situation: the classical action has no
dimensionful coupling, but quantum renormalization produces
\(\Lambda_{\rm QCD}\) and a nonzero trace anomaly, so the quantum theory is
not a conformal field theory.
\end{remark}

\section{The Weyl Map to the Cylinder}

Let \(x^\mu\) be Euclidean coordinates on \(\R^D\), and write
\[
  r=|x|,\qquad
  x^\mu=r n^\mu,\qquad
  n^\mu n_\mu=1.
\]
The flat metric is
\[
  \dd s^2
  =
  \dd r^2+r^2\dd\Omega_{D-1}^2,
\]
where \(\dd\Omega_{D-1}^2\) is the unit metric on \(S^{D-1}\).  Define
\[
  \tau=\log r .
\]
Then
\begin{equation}
  \frac{\dd s^2}{r^2}
  =
  \dd\tau^2+\dd\Omega_{D-1}^2.
  \label{eq:radial-cylinder-metric}
\end{equation}
Thus the punctured space \(\R^D\setminus\{0\}\) is Weyl equivalent to the
Euclidean cylinder
\[
  \R_\tau\times S^{D-1}.
\]
The origin corresponds to \(\tau=-\infty\), and infinity corresponds to
\(\tau=+\infty\).

The use of this Weyl map assumes that the flat-space conformal theory has
been extended to background metrics with a specified Weyl transformation law
for renormalized local operators.  In even dimensions a Weyl anomaly may add
local curvature terms to the generating functional.  On the punctured flat
space and cylinder comparison below, such terms affect contact and vacuum
normalization data; the separated-point operator map is fixed after a choice
of renormalized operator basis.

Figure~\ref{fig:radial-cylinder-map} displays the Weyl map from punctured flat
space to the cylinder, with radial distance becoming cylinder time
\(\tau=\log r\).

\begin{figure}[H]
  \centering
  \resizebox{0.94\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small] at (0,2.35) {flat space};
      \draw[thick] (-2.2,-1.3) rectangle (2.2,1.5);
      \draw[blue!60!black,thick] (0,0) circle (0.55);
      \draw[blue!60!black,thick] (0,0) circle (1.05);
      \draw[blue!60!black,thick] (0,0) circle (1.55);
      \draw[orange!85!black,thick,-{Latex[length=2mm]}] (0,0)--(1.2,0.9);
      \node[orange!85!black] at (0.78,0.70) {\(r\)};
      \fill[purple!75!black] (0,0) circle (2pt);
      \node[purple!75!black,below left] at (0,0) {\(0\)};
    \end{scope}

    \draw[-{Latex[length=2.5mm]},thick] (2.75,0.12)--(4.05,0.12);
    \node[align=center] at (3.40,0.63) {\(\tau=\log r\)\\Weyl rescale};

    \begin{scope}[shift={(6.3,0)}]
      \node[font=\small] at (0,2.35) {cylinder};
      \draw[thick] (-1.35,-1.45)--(-1.35,1.55);
      \draw[thick] (1.35,-1.45)--(1.35,1.55);
      \draw[thick] (-1.35,1.55) arc[start angle=180,end angle=360,x radius=1.35,y radius=0.28];
      \draw[thick,dashed] (-1.35,1.55) arc[start angle=180,end angle=0,x radius=1.35,y radius=0.28];
      \draw[thick] (-1.35,-1.45) arc[start angle=180,end angle=360,x radius=1.35,y radius=0.28];
      \draw[thick,dashed] (-1.35,-1.45) arc[start angle=180,end angle=0,x radius=1.35,y radius=0.28];
      \foreach \y in {-0.8,0.0,0.8} {
        \draw[blue!60!black,thick] (-1.35,\y) arc[start angle=180,end angle=360,x radius=1.35,y radius=0.28];
        \draw[blue!60!black,thick,dashed] (-1.35,\y) arc[start angle=180,end angle=0,x radius=1.35,y radius=0.28];
      }
      \draw[-{Latex[length=2mm]},orange!85!black,thick] (-1.75,-1.35)--(-1.75,1.45);
      \node[orange!85!black,left] at (-1.75,0.15) {\(\tau\)};
      \node at (0,-1.95) {\(S^{D-1}\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{Radial quantization identifies dilatations on \(\R^D\) with
  translations in cylinder time \(\tau=\log r\).}
  \label{fig:radial-cylinder-map}
\end{figure}

\section{Flat, Conformal, and Cylinder Vacua}
\label{sec:flat-conformal-cylinder-vacua}

Three vacuum vectors enter the radial construction, and their equality is a
statement with hypotheses rather than a convention.  The Lorentzian flat-space
vacuum will be denoted by \(\ket{\Omega_{\rm P}}\).  It is the normalized
Poincare-invariant vector in the Wightman Hilbert space on
\(\R^{1,D-1}\), with
\[
  P_\mu\ket{\Omega_{\rm P}}=0,
  \qquad
  M_{\mu\nu}\ket{\Omega_{\rm P}}=0,
\]
where \(P_\mu\) are the translation generators and \(M_{\mu\nu}\) are the
Lorentz generators.  A conformal vacuum is a normalized vector
\(\ket{\Omega_{\rm conf}}\) invariant under the connected conformal group, or
equivalently on a common invariant domain,
\[
  P_\mu\ket{\Omega_{\rm conf}}
  =
  M_{\mu\nu}\ket{\Omega_{\rm conf}}
  =
  D\ket{\Omega_{\rm conf}}
  =
  K_\mu\ket{\Omega_{\rm conf}}
  =
  0 .
\]
The cylinder vacuum \(\ket{\Omega_0}_{\rm cyl}\) is the normalized
zero-eigenvalue state of the radial Hamiltonian \(D_{\rm rad}\) in the Hilbert
space on \(S^{D-1}\):
\[
  D_{\rm rad}\ket{\Omega_0}_{\rm cyl}=0 .
\]
It is represented by the empty Euclidean ball, or equivalently by insertion of
the identity operator at the origin.

\begin{proposition}[Identification of the flat and conformal vacua]
\label{prop:poincare-vacuum-is-conformal-vacuum}
Assume that the connected conformal group, or its universal cover, is
implemented by strongly continuous unitaries on the Wightman Hilbert space on a
common dense invariant domain containing \(\ket{\Omega_{\rm P}}\).  Assume also
that the translation-invariant normalized vector is unique and that the
generators obey the conformal Lie algebra on this domain.  Then the flat
Poincare vacuum is invariant under the connected conformal group.  Thus
\(\ket{\Omega_{\rm conf}}=\ket{\Omega_{\rm P}}\), up to the irrelevant overall
phase of a normalized vector.
\end{proposition}

\begin{proof}
Let \(U_D(s)=\exp(\ii sD)\) be the unitary one-parameter group of dilatations.
The relation
\[
  U_D(s)^{-1}P_\mu U_D(s)=\ee^sP_\mu
\]
implies that \(U_D(s)\ket{\Omega_{\rm P}}\) is again annihilated by all
translations.  By uniqueness of the normalized translation-invariant vector,
\[
  U_D(s)\ket{\Omega_{\rm P}}=\ee^{\ii\alpha(s)}\ket{\Omega_{\rm P}} .
\]
Strong continuity makes \(\alpha(s)\) linear, so
\(D\ket{\Omega_{\rm P}}=d_0\ket{\Omega_{\rm P}}\) for some real number
\(d_0\).  The number \(d_0\) is forced to vanish by the conformal algebra:
because \(P_\nu\ket{\Omega_{\rm P}}=0\) and \(P_\nu\) is self-adjoint,
\[
  0
  =
  \bra{\Omega_{\rm P}}[P_\nu,K_\mu]\ket{\Omega_{\rm P}}
  =
  2\eta_{\mu\nu}d_0 .
\]
Thus \(D\ket{\Omega_{\rm P}}=0\), which is the representation-theoretic
normalization in which the identity operator has dimension zero.

Now use the conformal commutator
\[
  [P_\nu,K_\mu]=2\eta_{\mu\nu}D+2M_{\nu\mu}
\]
with Lorentzian metric \(\eta\).  Since \(P_\nu\), \(D\), and \(M_{\nu\mu}\)
annihilate \(\ket{\Omega_{\rm P}}\),
\[
  P_\nu K_\mu\ket{\Omega_{\rm P}}
  =
  [P_\nu,K_\mu]\ket{\Omega_{\rm P}}
  =
  0 .
\]
Therefore \(K_\mu\ket{\Omega_{\rm P}}\) is translation-invariant.  Uniqueness
of the translation-invariant vector gives
\[
  K_\mu\ket{\Omega_{\rm P}}=c_\mu\ket{\Omega_{\rm P}}
\]
for constants \(c_\mu\).  Lorentz invariance of \(\ket{\Omega_{\rm P}}\) and
the vector transformation law of \(K_\mu\) force \(c_\mu\) to be a Lorentz
invariant vector; hence \(c_\mu=0\).  Thus the vacuum is annihilated by
translations, Lorentz generators, dilatations, and special conformal
generators, which are the infinitesimal generators of the connected conformal
group.
\end{proof}

The relation between \(\ket{\Omega_{\rm conf}}\) and
\(\ket{\Omega_0}_{\rm cyl}\) is supplied by the Weyl map
\(\R^D\setminus\{0\}\to\R_\tau\times S^{D-1}\).  The Euclidean radial Hilbert
space is constructed from the same flat-space vacuum correlators by cutting
the Schwinger functions on spheres.  Under this construction the identity
operator at the origin creates the zero-energy state, so the common vacuum is
identified as
\[
  \ket{\Omega_{\rm P}}
  =
  \ket{\Omega_{\rm conf}}
  \longleftrightarrow
  \ket{\Omega_0}_{\rm cyl}.
\]
This identification fixes the notation used below: \(\ket{\vac}\) denotes this
common vacuum whenever the flat or cylinder realization is clear from context.

The literal cutting-and-sewing meaning of radial quantization is a
Kontsevich--Segal type Euclidean field-theoretic structure.  In that language,
a closed Euclidean hypersurface such as \(S^{D-1}\) is assigned a state space,
a Euclidean bordism such as an annulus or punctured ball is assigned a linear
amplitude between boundary state spaces, and gluing of bordisms is represented
by composition of amplitudes.  At this level the assigned state space should
not be assumed to be a Hilbert space.  In a general Euclidean field theory it
is more naturally a locally convex topological vector space, often a
Fr\'echet or nuclear space in analytic formulations, and the bordism
amplitudes are continuous linear maps or distributional pairings.

Only for a unitary CFT does the radial cutting construction produce the
Hilbert space used in the literal state--operator correspondence, and the
extra ingredient is the inversion pairing.  The map \(x\mapsto x/|x|^2\)
exchanges the inside and outside of a sphere, identifies the boundary sphere
with its reflected copy up to the conformal Weyl factor, and turns reflection
positivity of the doubled configuration into a positive semidefinite Hermitian
form.  Only after quotienting the null space and completing this form does one
obtain the radial Hilbert space used below.  Thus the hypotheses in this
chapter use the part of the Kontsevich--Segal cutting/gluing structure needed
for a CFT together with the inversion pairing: the state space on \(S^{D-1}\),
the positive radial inner product, the annular semigroup generated by
\(D_{\rm rad}\), and the sewing identities for cylinder slabs.  The complete
Kontsevich--Segal axiom system and its construction-status ledger are
developed in Chapter~\ref{ch:kontsevich-segal-functorial-qft}.  The use here
is the CFT specialization: literal radial quantization is first a
hypersurface-cutting and bordism-gluing construction, and in a CFT the
inversion pairing upgrades it to a Hilbert-space construction.

There remains a normalization convention.  If a cylinder Hamiltonian is defined
by integrating a chosen renormalized stress tensor on the curved cylinder, Weyl
anomaly terms and local curvature counterterms can add a scalar Casimir energy
to that operator.  The radial Hamiltonian \(D_{\rm rad}\) used in this chapter
is the shifted generator whose identity state has eigenvalue zero.  Equivalently
the vacuum functional is normalized so that \(\langle 1\rangle=1\) on flat
space and on the cylinder, with anomaly-dependent local curvature terms fixed
once and for all.  Separated-point operator correlators are then related by
\eqref{eq:cylinder-flat-correlator-equivalence}.

\section{Operators Under the Cylinder Map}

For a scalar primary \(\mathcal O(x)\) of dimension \(\Delta\), define the
cylinder operator
\begin{equation}
  \widetilde{\mathcal O}(\tau,n)
  =
  \ee^{\Delta\tau}\mathcal O(\ee^\tau n).
  \label{eq:scalar-cylinder-operator-map}
\end{equation}
This is the ordinary Weyl transformation of a scalar primary under
\(\dd s^2\mapsto r^{-2}\dd s^2\).

The spin factor is the local orthonormal-frame rotation induced by the same
map.  Work on a coordinate patch \(U\subset S^{D-1}\) with a smooth oriented
orthonormal tangent frame
\[
  E_A^\mu(n),
  \qquad
  A=1,\ldots,D-1,
  \qquad
  n_\mu E_A^\mu(n)=0,
  \qquad
  E_A^\mu(n)E_{B\mu}(n)=\delta_{AB}.
\]
The cylinder metric is \(\widehat g=r^{-2}\delta_{\mu\nu}\dd x^\mu\dd x^\nu\).
An orthonormal coframe for \(\widehat g\), pulled back to the punctured flat
space, is
\begin{equation}
  \widehat e^{\hat\tau}{}_\mu(x)=r^{-1}n_\mu,
  \qquad
  \widehat e^{\hat A}{}_\mu(x)=r^{-1}E_A{}_\mu(n).
  \label{eq:cylinder-local-orthonormal-coframe}
\end{equation}
Removing the common Weyl scale gives the \(SO(D)\) rotation from the flat
Cartesian orthonormal frame to the cylinder orthonormal frame,
\begin{equation}
  R^{\hat a}{}_\mu(n)
  :=
  r\,\widehat e^{\hat a}{}_\mu(x),
  \qquad
  R^{\hat\tau}{}_\mu(n)=n_\mu,
  \qquad
  R^{\hat A}{}_\mu(n)=E_A{}_\mu(n),
  \label{eq:cylinder-frame-rotation}
\end{equation}
so \(R^{\hat a}{}_\mu R^{\hat b\mu}=\delta^{\hat a\hat b}\).  On overlaps of
two patches the frames \(E_A\) differ by an \(SO(D-1)\) rotation; the formulas
below transform by the corresponding tensor or spin representation.

For a covariant vector primary \(V_\mu(x)\) of dimension \(\Delta\), the
cylinder orthonormal-frame components are
\begin{equation}
  \widetilde V_{\hat a}(\tau,n)
  =
  \ee^{\Delta\tau}
  R_{\hat a}{}^\mu(n)\,
  V_\mu(\ee^\tau n).
  \label{eq:vector-cylinder-operator-map}
\end{equation}
Equivalently,
\[
  V_\mu(\ee^\tau n)
  =
  \ee^{-\Delta\tau}
  R_\mu{}^{\hat a}(n)\,
  \widetilde V_{\hat a}(\tau,n).
\]
The associated spin-one state map is obtained by inserting the flat operator at
the origin:
\begin{equation}
  \ket{V_\mu}:=V_\mu(0)\ket{\vac},
  \qquad
  \lim_{\tau\to-\infty}
  \ee^{-\Delta\tau}\,
  \widetilde V_{\hat a}(\tau,n)\ket{\vac}
  =
  R_{\hat a}{}^\mu(n)\ket{V_\mu}.
  \label{eq:vector-primary-state-map}
\end{equation}
The direction dependence in the last expression is exactly the rotation between
the local cylinder frame at \(n\) and the fixed flat spin index at the origin.
For a primary in a representation \(\rho\) of \(SO(D)\), replace
\(R_{\hat a}{}^\mu\) in \eqref{eq:vector-cylinder-operator-map} by
\(\rho(R(n))\).

For a general renormalized local operator basis, the map can mix operators
with the same quantum numbers when local counterterms are changed.  The
statement used in radial quantization is that, after fixing this basis, the
correlation functions on the two Weyl-related backgrounds agree:
\begin{equation}
  \left\langle
    \mathcal O_1(x_1)\cdots\mathcal O_n(x_n)
  \right\rangle_{\R^D}
  =
  \left\langle
    \widetilde{\mathcal O}_1(\tau_1,n_1)\cdots
    \widetilde{\mathcal O}_n(\tau_n,n_n)
  \right\rangle_{\R\times S^{D-1}},
  \label{eq:cylinder-flat-correlator-equivalence}
\end{equation}
for separated points \(x_i=\ee^{\tau_i}n_i\), up to the specified local
curvature and contact terms.

\section{States from Local Operators}

Let
\[
  B_R=\{x\in\R^D:|x|<R\},
  \qquad
  \Sigma_R=\partial B_R=S_R^{D-1}.
\]
The state prepared by a local operator at the origin is defined, in the
radial reconstruction used in this chapter, by its radial pairings with the
dense states prepared by insertions inside \(B_R\).  If \(A\) is a finite
linear combination of renormalized local insertions supported in
\(B_R\setminus\{0\}\), and \(\Theta_R\) denotes reflection through
\(\Sigma_R\),
\[
  \Theta_R(x)=\frac{R^2x}{|x|^2},
\]
then the vector \(\ket{\mathcal O;R}\) is characterized on the dense
finite-energy domain by
\begin{equation}
  \left\langle A,\mathcal O;R\right\rangle_{\rm rad}
  =
  \left\langle
    \Theta_R A\,\mathcal O(0)
  \right\rangle_E .
  \label{eq:radial-state-exterior-pairing}
\end{equation}
Here \(\Theta_RA\) includes the Weyl factors and spin-frame rotations of the
reflected insertions.  Equation
\eqref{eq:radial-state-exterior-pairing} is the definition used for an
abstract CFT supplied by Schwinger distributions.  It uses the radial
reflection-positive pairing and the null quotient of
Hypothesis~\ref{hyp:radial-reconstruction-data}; it does not require a
continuum field-configuration measure.

\begin{definition}[Regulated ball preparation]
\label{def:regulated-ball-preparation}
When a Lagrangian representative is used as a mnemonic or as an actual
regularized construction, the ball preparation is the following finite
regulator family.  Map \(B_R\setminus B_\rho\) to the cylinder slab
\([\log\rho,\log R]\times S^{D-1}\).  Choose a radial time spacing \(a>0\),
a finite angular spectral cutoff \(L\) for the eigenmodes of the relevant
elliptic kinetic operator on \(S^{D-1}\), and a renormalized operator
representative \(\mathcal O_{\rho,a,L}\) supported in the inner cap
\(|x|\le\rho\).  The regulator parameter is
\[
  \varepsilon=(\rho,a,L),
  \qquad
  \rho\downarrow0,\quad a\downarrow0,\quad L\uparrow\infty .
  \label{eq:radial-ball-regulator-parameter}
\]
At fixed \(\varepsilon\), the boundary configuration space
\(\mathcal B_{\varepsilon,R}\) consists of the retained boundary modes on
\(\Sigma_R\).  The finite-dimensional trace map
\[
  \operatorname{Tr}_{\varepsilon,R}:
  \mathcal F_{\varepsilon,R}\longrightarrow \mathcal B_{\varepsilon,R}
\]
restricts a regulated bulk field configuration to those boundary modes.  The
finite-regulator measure or density on \(\mathcal F_{\varepsilon,R}\) factors
as
\begin{equation}
  d\nu_{\varepsilon,R}(\Phi)
  =
  d\nu_{\varepsilon,R}^{\rm bulk}
    (\Phi\mid\varphi_\varepsilon)\,
  d\lambda_{\varepsilon,R}^{\partial}(\varphi_\varepsilon),
  \qquad
  \varphi_\varepsilon=
  \operatorname{Tr}_{\varepsilon,R}\Phi .
  \label{eq:radial-ball-boundary-factorization}
\end{equation}
For even bosonic modes \(d\lambda_{\varepsilon,R}^{\partial}\) is an ordinary
finite-dimensional boundary measure.  For odd variables it is the
corresponding Berezin density, and in gauge theories it is supplied only
after the finite-regulator gauge-fixing or BV data have been specified.  The
conditioned ball wavefunctional is
\begin{equation}
  \Psi_{\mathcal O,R,\varepsilon}(\varphi_\varepsilon)
  =
  \int_{\operatorname{Tr}_{\varepsilon,R}\Phi=\varphi_\varepsilon}
  d\nu_{\varepsilon,R}^{\rm bulk}(\Phi\mid\varphi_\varepsilon)\,
  \exp\{-S_{\varepsilon,R}[\Phi]\}\,
  \mathcal O_{\rho,a,L} .
  \label{eq:regulated-ball-wavefunctional}
\end{equation}
The displayed conditioning is a finite-dimensional conditional integral over
interior regulated variables.  Gluing two regulated regions along
\(\Sigma_R\) means pairing their boundary wavefunctionals by integration
against \(d\lambda_{\varepsilon,R}^{\partial}\), with the Berezin or
gauge-fixed/BV analogue in the corresponding cases.

The continuum state represented by the regulated ball is the limit
\begin{equation}
  \ket{\mathcal O;R}
  =
  \lim_{\varepsilon\to0}
  \mathfrak R_{\varepsilon,R}
  \Psi_{\mathcal O,R,\varepsilon},
  \label{eq:ball-preparation-continuum-limit}
\end{equation}
whenever this limit exists in the radial Hilbert norm, or in the stated
distribution topology before Hilbert completion.  The map
\(\mathfrak R_{\varepsilon,R}\) embeds or identifies the finite-regulator
boundary state with the regulated radial state space.  In constructive
positive-measure examples, \eqref{eq:ball-preparation-continuum-limit} is
compatible with disintegration of the constructed Euclidean field over the
chosen boundary sigma-algebra or boundary Sobolev trace space.  In the general
CFT setting of this chapter, the mathematical state is the vector determined
by \eqref{eq:radial-state-exterior-pairing}, and
\eqref{eq:regulated-ball-wavefunctional} is a regulator-dependent
representative of that vector when such a representative has been supplied.
\end{definition}

After the Weyl map to the cylinder, changing \(R\) translates the state in
\(\tau\).  The state associated with \(\mathcal O\), with the radius
dependence absorbed into cylinder evolution, is denoted
\[
  \ket{\mathcal O}.
\]

\begin{definition}[Scaling operator classes and algebraic radial states]
\label{def:scaling-operator-classes}
A renormalized local operator class at the origin has definite scaling
dimension \(\Delta\) if, in all separated-point Euclidean correlators and with
the stated contact-term convention, its infinitesimal dilatation Ward identity
is
\[
  D_{\rm op}\mathcal O(0)=\Delta\,\mathcal O(0).
\]
Equivalently, after the Weyl map to the cylinder, the corresponding
origin-created state is an eigenvector of \(D_{\rm rad}\) with eigenvalue
\(\Delta\).  Let \(\mathcal V_{\rm loc}^{(\Delta)}\) be the quotient vector
space of such local scaling operators, including all spin components and
descendants of dimension \(\Delta\), modulo operators whose radial two-point
pairing with every local operator vanishes.

The algebraic local-operator space is the direct sum over scaling dimensions,
\[
  \mathcal V_{\rm loc}^{\rm alg}
  =
  \bigoplus_{\Delta\in\operatorname{Spec}_{\rm pp}(D_{\rm rad})}
  \mathcal V_{\rm loc}^{(\Delta)} ,
  \label{eq:algebraic-local-operator-space}
\]
where the symbol \(\oplus\) denotes finite linear combinations.  The matching
algebraic finite-energy subspace of the radial Hilbert space is
\[
  \mathcal H_{\rm alg}
  =
  \bigoplus_{\Delta\in\operatorname{Spec}_{\rm pp}(D_{\rm rad})}
  \ker(D_{\rm rad}-\Delta)
  \subset
  \mathcal H_{S^{D-1}} .
  \label{eq:algebraic-radial-state-space}
\]
Here \(\operatorname{Spec}_{\rm pp}\) denotes the point spectrum.  Under
Hypothesis~\ref{hyp:radial-reconstruction-data} the eigenspaces appearing in
\eqref{eq:algebraic-radial-state-space} have finite multiplicity and
\(\mathcal H_{\rm alg}\) is dense in \(\mathcal H_{S^{D-1}}\).
\end{definition}

\begin{definition}[Local radial boundary value of a state]
\label{def:local-radial-boundary-value-state}
Let \(\psi_a\), \(a=1,\ldots,m\), be a finite rotation multiplet of vectors in
\(\ker(D_{\rm rad}-\Delta)\).  The multiplet has local radial boundary-value
data if the following distributions are specified.

For every ordered list of renormalized local operators
\(\mathcal O_{I_1},\ldots,\mathcal O_{I_n}\), define on the configuration
space \(x_j\neq0\), \(x_i\neq x_j\), the punctured correlator
\[
  W_{\psi_a;I_1\ldots I_n}^{\circ}(x_1,\ldots,x_n)
  =
  \bra{\vac}
    \widetilde{\mathcal O}_{I_1}(\tau_1,n_1)\cdots
    \widetilde{\mathcal O}_{I_n}(\tau_n,n_n)
  \ket{\psi_a}_{\rm cyl},
  \qquad
  x_j=\ee^{\tau_j}n_j ,
  \label{eq:state-punctured-radial-correlator}
\]
first in every radial ordering chamber, and in other chambers by the
Euclidean permutation and contact prescriptions of the Schwinger hierarchy.
The required data are extensions
\[
  W_{\psi_a;I_1\ldots I_n}
  \in
  \mathcal D'\!\left((\R^D)^n\right)
  \label{eq:state-local-boundary-distribution}
\]
of these punctured distributions such that:
\begin{enumerate}
  \item their singular support is contained in the union of the ordinary
  partial diagonals \(x_i=x_j\) and the origin diagonals \(x_i=0\);
  \item their Euclidean covariance, spin transformation, and internal
  symmetry transformation laws are those of an additional insertion at the
  origin transforming in the same rotation and internal representation as the
  multiplet \(\psi_a\);
  \item their dilatation Ward identity has weight \(\Delta\) at the origin,
  namely scaling all \(x_j\) by \(\lambda>0\) multiplies the origin insertion
  by \(\lambda^{-\Delta}\), with the known anomalous-dimension matrix already
  diagonalized inside the finite-dimensional eigenspace;
  \item translation of the origin insertion is represented by the descendant
  multiplets \(P_{\mu_1}\cdots P_{\mu_k}\psi_a\), so differentiation with
  respect to the origin agrees, distributionally, with minus the sum of
  derivatives with respect to the exterior points;
  \item for every radius \(R\) and every finite inside-supported insertion
  state \(A\subset B_R\setminus\{0\}\), the exterior pairing obtained by
  reflecting \(A\) through \(\Sigma_R\) agrees with the Hilbert-space pairing,
  \[
    W_{\psi_a;\Theta_R A}
    =
    \langle A,\psi_a\rangle_{\rm rad}.
    \label{eq:local-boundary-value-pairing}
  \]
\end{enumerate}
The topology of the extension in
\eqref{eq:state-local-boundary-distribution} is the ordinary distribution
topology on compact subsets after choosing the local contact-term convention.
\end{definition}

\begin{theorem}[Operator from a definite-energy radial state]
\label{thm:operator-from-state-radial-boundary-value}
Assume Hypothesis~\ref{hyp:radial-reconstruction-data}.  Let
\(\psi_a\in\ker(D_{\rm rad}-\Delta)\) be a finite rotation multiplet with
local radial boundary-value data in the sense of
Definition~\ref{def:local-radial-boundary-value-state}.  Then there is a
unique class
\[
  [\mathcal O_{\psi_a}]
  \in
  \mathcal V_{\rm loc}^{(\Delta)}
\]
such that, for every exterior product of local operators, its Euclidean
correlators are the distributions
\eqref{eq:state-local-boundary-distribution}.  Its radial state is precisely
\(\psi_a\):
\[
  \ket{\mathcal O_{\psi_a}}=\psi_a .
  \label{eq:operator-from-state-radial-state}
\]
Two representatives that give the same radial state differ by a null local
operator, namely by an element whose radial two-point pairing with every local
operator vanishes.

Conversely, the radial-completeness clause of
Hypothesis~\ref{hyp:radial-reconstruction-data} is exactly the assertion that
every vector in each eigenspace \(\ker(D_{\rm rad}-\Delta)\) admits such local
radial boundary-value data.  Under that clause, the operator-from-state map is
surjective separately on every eigenspace.
\end{theorem}

\begin{proof}
The extensions
\eqref{eq:state-local-boundary-distribution} define the correlators of an
additional distributional insertion supported at the origin.  The singular
support condition and the stated contact prescriptions give locality at all
separated points and a prescribed extension across origin diagonals.  The
covariance and scaling Ward identities place this insertion in the
\(\Delta\)-eigenspace of the local dilatation action, with the specified spin
and internal quantum numbers, and the translation condition identifies its
descendants with the radial states \(P_{\mu_1}\cdots P_{\mu_k}\psi_a\).
Thus the data define a class
\([\mathcal O_{\psi_a}]\in\mathcal V_{\rm loc}^{(\Delta)}\).

The radial state of this class is computed by pairing it with the dense set of
inside-supported insertion states.  For every such \(A\), the defining
pairing \eqref{eq:radial-state-exterior-pairing} and the boundary-value
condition \eqref{eq:local-boundary-value-pairing} give
\[
  \langle A,\mathcal O_{\psi_a};R\rangle_{\rm rad}
  =
  \langle A,\psi_a\rangle_{\rm rad}.
\]
Since these \(A\)'s are dense after the radial null quotient, the two radial
vectors agree.  If two local insertions have the same radial state, their
difference pairs to zero with every \(A\), hence with every local operator by
radial reflection and density; this is precisely the null equivalence used in
the definition of \(\mathcal V_{\rm loc}^{(\Delta)}\).

The final assertion is a restatement of radial completeness with the
existence condition made explicit: completeness supplies, for every
finite-energy eigenvector, a local scaling field whose correlators furnish the
boundary-value distributions above.
\end{proof}

\begin{construction}[cylinder sewing and Heisenberg operators]
\label{constr:cylinder-sewing-heisenberg}
Let
\[
  \Sigma_\tau=\{\tau\}\times S^{D-1}
\]
be a constant-cylinder-time slice, and write
\[
  H_{\rm cyl}=D_{\rm rad}
\]
for the nonnegative self-adjoint generator of translations in \(\tau\).
For a cylinder local operator \(X(n)\) placed on the slice \(\Sigma_0\), its
Euclidean Heisenberg translate is the operator on the common finite-energy
domain
\begin{equation}
  X(\tau,n)=
  \ee^{\tau H_{\rm cyl}}X(0,n)\ee^{-\tau H_{\rm cyl}} .
  \label{eq:cylinder-heisenberg-operator}
\end{equation}
For separated insertions
\[
  \tau_+>\tau_1>\tau_2>\cdots>\tau_n>\tau_- ,
\]
and boundary states \(v_-\) on \(\Sigma_{\tau_-}\) and \(v_+\) on
\(\Sigma_{\tau_+}\), the cylinder amplitude obtained by cutting along all
intermediate slices is
\begin{align}
  &\mathcal A_{\tau_+,\tau_-}
  (v_+;X_1,\ldots,X_n;v_-)
  \nonumber\\
  &\quad =
  \left\langle
    v_+,
    \ee^{-(\tau_+-\tau_1)H_{\rm cyl}}
    X_1(0,n_1)
    \ee^{-(\tau_1-\tau_2)H_{\rm cyl}}
    \cdots
    X_n(0,n_n)
    \ee^{-(\tau_n-\tau_-)H_{\rm cyl}}
    v_-
  \right\rangle .
  \label{eq:cylinder-slab-sewing}
\end{align}
Equation \eqref{eq:cylinder-slab-sewing} is the Hilbert-space form of
Euclidean sewing.  In a functional-integral presentation it is obtained by
integrating over the boundary field data on each cut \(\Sigma_{\tau_j}\); in
the radial reconstruction presentation it is the composition law for the
positive-time semigroup on the null quotient.  The two descriptions give the
same sesquilinear form because both are defined by the same Schwinger
correlation hierarchy.

Taking \(v_+=\vac\) and using
\[
  \bra{\vac}\ee^{-sH_{\rm cyl}}=\bra{\vac},
  \qquad s\ge0,
\]
turns \eqref{eq:cylinder-slab-sewing} into the Euclidean Heisenberg matrix
element
\begin{equation}
  \bra{\vac}
  X_1(\tau_1,n_1)\cdots X_n(\tau_n,n_n)
  \ket{v}
  =
  \mathcal A_{\tau_+,\tau_-}
  \bigl(\vac;X_1,\ldots,X_n;\ee^{-\tau_-H_{\rm cyl}}v\bigr),
  \label{eq:cylinder-sewing-heisenberg-matrix-element}
\end{equation}
for every finite-energy vector \(v\) and every lower cut
\(\tau_-<\tau_n\).  The right side is independent of the auxiliary cuts after
the boundary state is transported as displayed.  Thus the cylinder correlation
function with a state prepared in the far past is computed by an ordered
product of Heisenberg cylinder operators.
\end{construction}

\begin{theorem}[State--operator correspondence]
\label{thm:state-operator-correspondence}
Assume the radial reconstruction data of
Hypothesis~\ref{hyp:radial-reconstruction-data}.  For every
\(\Delta\in\operatorname{Spec}_{\rm pp}(D_{\rm rad})\), insertion at the
origin defines an isometric linear isomorphism between local scaling operators
of dimension \(\Delta\) and cylinder energy eigenstates of energy \(\Delta\):
\begin{equation}
  \mathsf S_\Delta:
  \mathcal V_{\rm loc}^{(\Delta)}
  \xrightarrow{\;\sim\;}
  \ker(D_{\rm rad}-\Delta),
  \qquad
  \mathcal O\longmapsto\ket{\mathcal O}.
  \label{eq:state-operator-map-degree-delta}
\end{equation}
Consequently finite linear combinations of definite-dimension local operators
give an isometric algebraic isomorphism
\[
  \mathsf S_{\rm alg}:
  \mathcal V_{\rm loc}^{\rm alg}
  \xrightarrow{\;\sim\;}
  \mathcal H_{\rm alg}.
\]
The BPZ, or radial, completion of
\(\mathcal V_{\rm loc}^{\rm alg}\) is identified with
\(\mathcal H_{S^{D-1}}\), because \(\mathcal H_{\rm alg}\) is dense.  A
generic vector in this completed Hilbert space is a state on the sphere; it is
represented by a local operator precisely when it lies in the algebraic
finite-energy subspace with the local boundary-value property above.  For a
definite-dimension operator and for insertions ordered by cylinder time
\(\tau_1>\tau_2>\cdots>\tau_n\),
\begin{equation}
  \left\langle
  \mathcal O(0)\mathcal O_1(x_1)\cdots\mathcal O_n(x_n)
  \right\rangle_{\R^D}
  =
  \bra{\vac}
  \widetilde{\mathcal O}_1(\tau_1,n_1)\cdots
  \widetilde{\mathcal O}_n(\tau_n,n_n)
  \ket{\mathcal O}_{\rm cyl}.
  \label{eq:state-operator-correlation-formula}
\end{equation}
\end{theorem}

\begin{proof}
Radial reflection positivity supplies the positive semidefinite inner product
on local insertions inside a sphere.  Quotienting its null space makes the map
\(\mathcal O\mapsto\ket{\mathcal O}\) injective, and the defining pairing
\eqref{eq:radial-state-exterior-pairing} identifies its norm with the radial
two-point pairing; hence the map is isometric.  In a regulated Lagrangian
representative this equality is the gluing identity obtained from
\eqref{eq:radial-ball-boundary-factorization}, followed by the continuum
limit \eqref{eq:ball-preparation-continuum-limit}.  If
\(\mathcal O\in\mathcal V_{\rm loc}^{(\Delta)}\), the dilatation Ward identity
for \(\mathcal O(0)\) becomes the cylinder Hamiltonian equation
\[
  D_{\rm rad}\ket{\mathcal O}=\Delta\ket{\mathcal O}.
\]
Thus \(\mathsf S_\Delta\) maps
\(\mathcal V_{\rm loc}^{(\Delta)}\) isometrically into
\(\ker(D_{\rm rad}-\Delta)\).  Theorem
\ref{thm:operator-from-state-radial-boundary-value}, together with the
radial-completeness clause of
Hypothesis~\ref{hyp:radial-reconstruction-data}, gives the reverse inclusion:
each vector in \(\ker(D_{\rm rad}-\Delta)\) is represented by a local scaling
operator of the same dimension, unique modulo null operators.  This proves
\eqref{eq:state-operator-map-degree-delta}.  Taking finite direct sums gives
the algebraic isomorphism.  Since \(D_{\rm rad}\) is nonnegative self-adjoint,
the spectral projections \(P_{[0,E]}(D_{\rm rad})\) converge strongly to the
identity as \(E\to\infty\).  Hence \(\mathcal H_{\rm alg}\) is dense in
\(\mathcal H_{S^{D-1}}\), and the isometry extends by completion to the radial
Hilbert space.  This completion is a completion of states, not an assertion
that every Hilbert-space vector is itself a local operator.

It remains to derive the correlation formula.  Put
\[
  X_j(0,n_j)=\widetilde{\mathcal O}_j(0,n_j),
  \qquad
  X_j(\tau_j,n_j)=
  \widetilde{\mathcal O}_j(\tau_j,n_j),
\]
where the second equality is the Heisenberg translation
\eqref{eq:cylinder-heisenberg-operator}.  Choose cuts
\(\tau_+>\tau_1\) and \(\tau_-<\tau_n\).  By
\eqref{eq:cylinder-sewing-heisenberg-matrix-element} with
\(v=\ket{\mathcal O}_{\rm cyl}\),
\[
  \bra{\vac}
  \widetilde{\mathcal O}_1(\tau_1,n_1)\cdots
  \widetilde{\mathcal O}_n(\tau_n,n_n)
  \ket{\mathcal O}_{\rm cyl}
\]
is the cylinder Schwinger functional on the punctured cylinder with:
\begin{itemize}
\item the incoming boundary state prepared by the local insertion
  \(\mathcal O(0)\), transported to the lower cut;
\item insertions \(X_n,\ldots,X_1\) placed in the slabs ordered by increasing
  cylinder time; and
\item the outgoing boundary projected to the vacuum by the limit
  \(\tau_+\to+\infty\).
\end{itemize}
The vacuum projection exists because \(H_{\rm cyl}\ge0\) and the zero-energy
subspace is \(\mathbb C\vac\).  The sewing identity is unchanged if the
auxiliary cuts are moved without crossing an insertion.

Finally apply the Weyl map \(x_j=\ee^{\tau_j}n_j\).  The operator definition
\eqref{eq:scalar-cylinder-operator-map}, with the corresponding spin-frame
factor for spinning operators, identifies the cylinder Schwinger functional
at separated points with the flat-space radially ordered correlator.  Since
\(\tau_1>\cdots>\tau_n\), radial ordering places
\(\mathcal O_1(x_1),\ldots,\mathcal O_n(x_n)\) outside the origin insertion in
the displayed order.  This gives
\eqref{eq:state-operator-correlation-formula}.

\end{proof}

The state \(\ket{\vac}\) is the ground state on \(S^{D-1}\).

The vacuum state corresponds to the identity operator:
\[
  \mathbf 1\longleftrightarrow\ket{\vac}.
\]
The algebraic local-operator space is dense in the BPZ/radial Hilbert norm but
is not closed.  Accordingly, the product of two separated local insertions
inside a sphere defines a state on that sphere, not a new local operator at
the origin.  The operator product expansion is the statement that this state
is the BPZ/radial-norm limit of finite sums of local scaling operators at the
origin, organized by \(D_{\rm rad}\)-eigenspaces.  The density of
\(\mathcal V_{\rm loc}^{\rm alg}\) in this norm is the functional-analytic
input that permits separated products to be represented by convergent local
operator partial sums after radial ordering.

Figure~\ref{fig:state-from-origin-cylinder} records the precise
state-from-operator construction: inserting a local scaling operator at the
origin gives an energy eigenstate on the sphere, and separated products give
states approximated by local-operator partial sums.

\begin{figure}[H]
  \centering
  \resizebox{0.95\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      insertion/.style={circle,fill=violet!75!black,inner sep=2pt},
      sphere/.style={green!45!black,line width=0.9pt},
      arrow/.style={-{Latex[length=2mm]},thick}
    }
    \begin{scope}
      \node[font=\small] at (0,2.20) {punctured flat space};
      \draw[thick] (-2.55,-1.55) rectangle (2.55,1.70);
      \draw[sphere] (0,0) ellipse (1.30 and 0.48);
      \draw[sphere] (0,0) circle (1.00);
      \node[insertion] at (0,0) {};
      \node[violet!75!black,above] at (0,0.12) {\(\mathcal O(0)\)};
      \node[insertion] at (1.75,-0.92) {};
      \node[violet!75!black,below left] at (1.75,-0.92) {\(\mathcal O'(x)\)};
      \draw[blue!65!black,arrow] (0,0)--(0,1.00);
      \draw[blue!65!black,arrow] (0,0)--(1.15,0.40);
      \draw[blue!65!black,arrow] (0,0)--(-0.90,-0.35);
      \node[blue!65!black] at (1.10,0.82) {radial time};
    \end{scope}

    \draw[arrow] (3.05,0.10)--(4.35,0.10);
    \node[align=center] at (3.70,0.62) {\(r=\ee^\tau\)\\and Weyl rescale};

    \begin{scope}[shift={(6.95,0)}]
      \node[font=\small] at (0,2.20) {cylinder};
      \draw[thick] (-1.45,-1.45)--(-1.45,1.55);
      \draw[thick] (1.45,-1.45)--(1.45,1.55);
      \draw[thick] (-1.45,1.55) arc[start angle=180,end angle=360,x radius=1.45,y radius=0.30];
      \draw[thick,dashed] (-1.45,1.55) arc[start angle=180,end angle=0,x radius=1.45,y radius=0.30];
      \draw[thick] (-1.45,-1.45) arc[start angle=180,end angle=360,x radius=1.45,y radius=0.30];
      \draw[thick,dashed] (-1.45,-1.45) arc[start angle=180,end angle=0,x radius=1.45,y radius=0.30];
      \draw[sphere] (-1.45,-0.35) arc[start angle=180,end angle=360,x radius=1.45,y radius=0.30];
      \draw[sphere,dashed] (-1.45,-0.35) arc[start angle=180,end angle=0,x radius=1.45,y radius=0.30];
      \draw[blue!65!black,arrow] (-1.95,-1.35)--(-1.95,1.35);
      \node[blue!65!black,left] at (-1.95,0.10) {\(\tau\)};
      \node[violet!75!black] at (-0.85,-1.22) {\(\ket{\mathcal O}\)};
      \node[violet!75!black] at (0.95,0.92) {\(\widetilde{\mathcal O}'(\tau,n)\)};
      \node at (0,-1.95) {spatial sphere \(S^{D-1}\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The singular endpoint \(r=0\) of the Weyl map becomes
  \(\tau=-\infty\) on the cylinder.  A local insertion at the origin therefore
  prepares a state on the spatial sphere.}
  \label{fig:state-from-origin-cylinder}
\end{figure}

\section{Dilatation as the Cylinder Hamiltonian}

The dilatation vector field on \(\R^D\) is
\[
  x^\mu\partial_\mu=r\partial_r=\partial_\tau.
\]
Therefore the dilatation charge maps to the Hamiltonian generating
translations in cylinder time.  Let \(\widehat H_{\mathrm{cyl}}\) be defined
by Euclidean evolution
\[
  \ket{\psi(\tau+a)}
  =
  \ee^{-a\widehat H_{\mathrm{cyl}}}\ket{\psi(\tau)} .
\]
In radial-quantization conventions the factors of \(\ii\) in the Lorentzian
Hermitian conformal charges are absorbed, and
\begin{equation}
  \widehat H_{\mathrm{cyl}}=\widehat D_{\rm rad}.
  \label{eq:dilatation-cylinder-hamiltonian}
\end{equation}
The eigenvalue of \(\widehat H_{\mathrm{cyl}}\) is the scaling dimension.

The relation between the Hermitian Lorentzian convention and the radial
Euclidean convention must be tracked at the level of the complexified
conformal algebra.  Let \(\widehat X_{\rm L}\) denote the Hermitian
Lorentzian charges in the convention of
Chapter~\ref{ch:volume-iii-conformal-killing-vectors}, so for example
\[
  [\widehat D_{\rm L},\widehat P^{\rm L}_\mu]
  =
  \ii\widehat P^{\rm L}_\mu,
  \qquad
  [\widehat P^{\rm L}_\mu,\widehat K^{\rm L}_\nu]
  =
  2\ii
  (\delta_{\mu\nu}\widehat D_{\rm L}-\widehat J^{\rm L}_{\mu\nu}).
\]
The radial descendant algebra is the real form adapted to the cylinder
Hamiltonian and to the inversion pairing.  With the subscript ``rad'' used
only in this paragraph, the conversion is
\begin{equation}
  \widehat D_{\rm rad}=-\ii\widehat D_{\rm L},
  \qquad
  \widehat J^{\rm rad}_{\mu\nu}=-\ii\widehat J^{\rm L}_{\mu\nu},
  \qquad
  \widehat P^{\rm rad}_{\mu}=\ii\widehat P^{\rm L}_{\mu},
  \qquad
  \widehat K^{\rm rad}_{\mu}=-\ii\widehat K^{\rm L}_{\mu}.
  \label{eq:lorentzian-to-radial-generator-map}
\end{equation}
The opposite signs for \(\widehat P_\mu\) and \(\widehat K_\mu\) reflect
their radial roles: translations create descendants and raise cylinder
energy, while special conformal transformations annihilate primaries and
lower cylinder energy.  Applying \eqref{eq:lorentzian-to-radial-generator-map}
gives
\[
  [\widehat D_{\rm rad},\widehat P^{\rm rad}_\mu]
  =
  \widehat P^{\rm rad}_\mu,
  \qquad
  [\widehat D_{\rm rad},\widehat K^{\rm rad}_\mu]
  =
  -\widehat K^{\rm rad}_\mu,
\]
and
\begin{equation}
  [\widehat P^{\rm rad}_\mu,\widehat K^{\rm rad}_\nu]
  =
  -2(\delta_{\mu\nu}\widehat D_{\rm rad}
      -\widehat J^{\rm rad}_{\mu\nu}),
  \qquad
  [\widehat K^{\rm rad}_\mu,\widehat P^{\rm rad}_\nu]
  =
  2\delta_{\mu\nu}\widehat D_{\rm rad}
  -2\widehat J^{\rm rad}_{\nu\mu}.
  \label{eq:radial-pk-sign-from-lorentzian}
\end{equation}
The radial chapters suppress the subscript after this point.  Thus the
transition from Hermitian Lorentzian charges to the radial algebra is not a
rule that deletes every visible factor of \(\ii\); it is the complex change of
real form displayed in \eqref{eq:lorentzian-to-radial-generator-map}.
Positivity statements then use the Hilbert-space adjoint
\(\widehat P_\mu^\dagger=\widehat K_\mu\) supplied by radial reflection.

If \(\mathcal O\) has scaling dimension \(\Delta\), then
\[
  \widehat H_{\mathrm{cyl}}\ket{\mathcal O}
  =
  \Delta\ket{\mathcal O}.
\]
Thus the spectrum of scaling dimensions in flat space is the energy spectrum
of the theory on the spatial sphere.  Unitarity implies \(\Delta\ge0\) for
states above the vacuum in a unitary CFT.

\section{Primary and Descendant States}

A local operator at the origin is called a conformal primary if the associated
state is annihilated by all special conformal generators:
\begin{equation}
  \widehat K_\mu\ket{\mathcal O}=0 .
  \label{eq:primary-state-condition}
\end{equation}
If \(\mathcal O_\alpha\) transforms in a finite-dimensional spin
representation with Hermitian matrices \((S_{\mu\nu})_\alpha{}^\beta\), so
that a finite rotation near the identity is represented by
\(\exp(-\frac{\ii}{2}\omega^{\mu\nu}S_{\mu\nu})\), then a primary state obeys
\[
  \widehat D_{\rm rad}\ket{\mathcal O_\alpha}
  =
  \Delta\ket{\mathcal O_\alpha},
  \qquad
  \widehat J_{\mu\nu}\ket{\mathcal O_\alpha}
  =
  -(S_{\mu\nu})_\alpha{}^\beta\ket{\mathcal O_\beta},
  \qquad
  \widehat K_\mu\ket{\mathcal O_\alpha}=0 .
\]
Translations generate descendants:
\[
  \widehat P_{\mu_1}\cdots\widehat P_{\mu_n}\ket{\mathcal O}
  \longleftrightarrow
  \partial_{\mu_1}\cdots\partial_{\mu_n}\mathcal O(0).
\]
The commutator
\[
  [\widehat D_{\rm rad},\widehat P_\mu]=\widehat P_\mu
\]
shows that each derivative raises the scaling dimension by one.
Before quotienting null relations, these descendant states form the
conformal generalized Verma module generated by the primary datum
\((\Delta,\rho)\); the algebraic definition is given in
Definition~\ref{def:conformal-generalized-verma-module}.  The Hilbert-space
primary multiplet is the quotient by the radical of the radial inner product,
and a shortening equation is precisely the vanishing of a descendant in this
quotient.

Figure~\ref{fig:primary-descendant-multiplet} displays the conformal multiplet
structure used in this quotient: a primary state generates descendants by
successive actions of \(\widehat P_\mu\), with null descendants removed in the
Hilbert completion.

\begin{figure}[H]
  \centering
  \resizebox{0.82\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,2.65) {conformal multiplet};
    \node[draw=black,rounded corners,inner sep=7pt] (p) at (0,1.75)
      {primary \(\ket{\mathcal O}\)};
    \node[draw=black,rounded corners,inner sep=7pt] (d1) at (-2.4,0.55)
      {\(\widehat P_\mu\ket{\mathcal O}\)};
    \node[draw=black,rounded corners,inner sep=7pt] (d2) at (2.4,0.55)
      {\(\widehat P_\nu\ket{\mathcal O}\)};
    \node[draw=black,rounded corners,inner sep=7pt] (d3) at (0,-0.75)
      {\(\widehat P_\mu\widehat P_\nu\ket{\mathcal O}\)};
    \draw[-{Latex[length=2mm]},thick] (p)--(d1);
    \draw[-{Latex[length=2mm]},thick] (p)--(d2);
    \draw[-{Latex[length=2mm]},thick] (d1)--(d3);
    \draw[-{Latex[length=2mm]},thick] (d2)--(d3);
    \node[blue!65!black] at (-3.8,1.15) {\(+1\) in \(\Delta\)};
    \node[orange!85!black] at (3.85,1.15) {\(\widehat K_\mu\) lowers};
    \node[orange!85!black,align=center] at (0,-1.65)
      {the primary is the lowest-energy state\\in the multiplet};
  \end{tikzpicture}
  }
  \caption{A conformal multiplet is generated from a primary by translations.
  On the cylinder, descendants have energies shifted upward by integers.}
  \label{fig:primary-descendant-multiplet}
\end{figure}

\section{Reflection Positivity and Conjugation}

Radial quantization gives a natural conjugation.  In Euclidean flat space,
reflection through the unit sphere is inversion:
\[
  x^\mu\mapsto \frac{x^\mu}{x^2}.
\]
On the cylinder this is time reflection \(\tau\mapsto-\tau\).  For a scalar
operator of dimension \(\Delta\), the conjugate insertion at infinity is
\[
  \mathcal O^\dagger(\infty)
  =
  \lim_{r\to\infty}r^{2\Delta}\mathcal O(r n)
\]
when the limit is independent of \(n\).  The two-point function becomes the
inner product
\[
  \braket{\mathcal O}{\mathcal O}
  =
  \left\langle
  \mathcal O^\dagger(\infty)\mathcal O(0)
  \right\rangle .
\]
For unitary theories this inner product is positive semidefinite and becomes
positive definite after null states are quotiented.  This positivity is the
source of unitarity bounds, developed after primary transformation laws and
two-point functions have been fixed.

\section{Adjoints of the Conformal Generators}

In radial quantization, Hermitian conjugation is a Hilbert-space statement in
the reflection-positive completion of
Hypothesis~\ref{hyp:radial-reconstruction-data}.  Let
\(\mathcal D_{\rm loc}\subset\mathcal H_{S^{D-1}}\) be the dense domain of
finite linear combinations of local finite-energy states.  For a scalar
primary, the BPZ bra associated with the ket
\[
  \ket{\mathcal O}
  =
  \lim_{\tau\to-\infty}
  \ee^{-\Delta\tau}\widetilde{\mathcal O}(\tau,n)\ket{\vac}
\]
is
\begin{equation}
  \langle\!\langle\mathcal O|
  =
  \lim_{\tau\to+\infty}
  \ee^{\Delta\tau}
  \bra{\vac}\widetilde{\mathcal O}^{\,\Theta_{\rm int}}(\tau,n),
  \label{eq:bpz-bra-radial-quantization-ch4}
\end{equation}
where \(\Theta_{\rm int}\) denotes the ordinary internal adjunction; for a
spinning primary one also includes the spin-frame rotation specified by the
radial reflection map in Hypothesis~\ref{hyp:radial-reconstruction-data}.
Equivalently, in flat coordinates the bra is represented by the insertion
obtained from \(x\mapsto x/x^2\), together with the Weyl factor and spin
transport.  Formula \eqref{eq:bpz-bra-radial-quantization-ch4} is the
Hilbert-space version of the cylinder limit used later in
\eqref{eq:bpz-bra-from-cylinder-limit}.

The adjoint of an unbounded conformal generator is first defined on
\(\mathcal D_{\rm loc}\).  Let \(T_a\) be the finite translation of inside
insertions by a small vector \(a\), and let
\begin{equation}
  C_a:=\vartheta\circ T_a\circ\vartheta,
  \qquad
  \vartheta(x)=\frac{x}{|x|^2},
  \label{eq:sct-as-reflected-translation}
\end{equation}
be the corresponding finite special conformal transformation of the reflected
insertions.  The equality of Schwinger functions after applying the same
conformal transformation to all separated insertions gives, for
\(A,B\in\mathcal D_{\rm loc}\) and \(a\) small enough that no support crosses the
reflection sphere,
\begin{equation}
  \langle A,T_a B\rangle_{\rm rad}
  =
  \langle C_a A,B\rangle_{\rm rad}.
  \label{eq:bpz-translation-special-conformal-finite}
\end{equation}
This is the step supplied by the BPZ bra construction: the transformation of
the ket insertion is moved to the bra by radial reflection, and inversion
turns the translated reflected insertion into a special-conformal one.  With
the radial generator convention in which translations create descendants,
\[
  \left.\frac{\partial}{\partial a^\mu}T_a B\right|_{a=0}
  =
  \widehat P_\mu B,
  \qquad
  \left.\frac{\partial}{\partial a^\mu}C_a A\right|_{a=0}
  =
  \widehat K_\mu A .
\]
Differentiating
\eqref{eq:bpz-translation-special-conformal-finite} at \(a=0\) gives
\begin{equation}
  \langle A,\widehat P_\mu B\rangle_{\rm rad}
  =
  \langle \widehat K_\mu A,B\rangle_{\rm rad},
  \qquad A,B\in\mathcal D_{\rm loc}.
  \label{eq:bpz-p-adjoint-k-domain}
\end{equation}
Thus \(\widehat K_\mu\) is the Hilbert-space adjoint of \(\widehat P_\mu\) on
the local finite-energy domain; after quotienting null states and taking the
closure this is the adjoint relation used in the descendant Gram matrices.

Dilatations are translations in cylinder time, so
\[
  \widehat D^\dagger=\widehat D
\]
in the cylinder Hilbert-space convention.  Rotations are unitary symmetries
of the spatial sphere:
\[
  \widehat J_{\mu\nu}^\dagger=\widehat J_{\mu\nu}.
\]
Equation \eqref{eq:bpz-p-adjoint-k-domain} is therefore summarized as
\begin{equation}
  \widehat P_\mu^\dagger=\widehat K_\mu,
  \qquad
  \widehat K_\mu^\dagger=\widehat P_\mu .
  \label{eq:radial-adjoint-p-k}
\end{equation}

Equation \eqref{eq:radial-adjoint-p-k} is the algebraic input behind
unitarity bounds.  If \(\ket{\mathcal O}\) is a primary state, then the norm
of a first descendant is
\[
  \|\widehat P_\mu\ket{\mathcal O}\|^2
  =
  \bra{\mathcal O}\widehat K_\mu\widehat P_\mu\ket{\mathcal O},
\]
with summation over \(\mu\) when a rotational scalar norm is taken.  Commuting
\(\widehat K_\mu\) through \(\widehat P_\nu\) and using
\(\widehat K_\mu\ket{\mathcal O}=0\) expresses the result in terms of
\(\Delta\) and the spin representation of the primary.  Positivity of all
descendant norms is therefore a representation-theoretic constraint on
allowed primary data.

\section{The Free Scalar Check}

Let \(\phi\) be a free massless scalar primary with
\[
  \Delta_\phi=\frac{D-2}{2}.
\]
The flat action is the \(R=0\) representative of the Weyl-covariant scalar
action
\begin{equation}
  S[\phi;g]
  =
  \frac12\int \dd^D x\,\sqrt g\,
  \left(
    g^{\mu\nu}\nabla_\mu\phi\nabla_\nu\phi
    +\xi_\ast R[g]\phi^2
  \right),
  \qquad
  \xi_\ast=\frac{D-2}{4(D-1)}.
  \label{eq:conformal-scalar-action-radial-check}
\end{equation}
Thus ``massless in flat space'' means that the curvature scalar vanishes in
the flat representative; it does not mean that the Weyl-covariant
representative lacks the curvature coupling.  By the Weyl transformation
already described in Chapter~\ref{ch:volume-iii-stress-tensor-weyl}, the same
theory on the cylinder has \(R_{\rm cyl}>0\), so the
\(\xi_\ast R\phi^2\) term is activated and becomes the cylinder mass term.  If
\[
  \widetilde\phi(\tau,n)=\ee^{\Delta_\phi\tau}\phi(\ee^\tau n),
\]
then
\begin{equation}
  S_{\rm cyl}[\widetilde\phi]
  =
  \frac12\int_{\R\times S^{D-1}}\dd\tau\,\dd\Omega_{D-1}
  \left[
    (\partial_\tau\widetilde\phi)^2
    +
    \nabla_{S^{D-1}}\widetilde\phi\cdot
    \nabla_{S^{D-1}}\widetilde\phi
    +
    \xi_\ast R_{\rm cyl}\widetilde\phi^2
  \right].
  \label{eq:free-scalar-cylinder-action}
\end{equation}
The cylinder metric is
\[
  \dd s_{\rm cyl}^2=\dd\tau^2+\dd\Omega_{D-1}^2,
\]
where \(\dd\Omega_{D-1}^2=\gamma_{ab}\dd\theta^a\dd\theta^b\) is the unit
round metric on \(S^{D-1}\).  The Levi-Civita connection of this product
metric splits.  Thus
\[
  R_{\tau\tau}^{\rm cyl}=0,
  \qquad
  R_{ab}^{\rm cyl}=R_{ab}^{S^{D-1}}=(D-2)\gamma_{ab},
\]
and contraction with \(g_{\rm cyl}^{\tau\tau}=1\) and
\(g_{\rm cyl}^{ab}=\gamma^{ab}\) gives
\[
  R_{\rm cyl}
  =
  g_{\rm cyl}^{MN}R_{MN}^{\rm cyl}
  =
  \gamma^{ab}(D-2)\gamma_{ab}
  =
  (D-1)(D-2).
\]
The conformal mass term is
\[
  m_{\mathrm{cyl}}^2=\xi_\ast R_{\rm cyl}=\frac{(D-2)^2}{4}.
\]
For a one-particle mode on \(S^{D-1}\) with angular momentum \(\ell\),
\[
  -\nabla^2_{S^{D-1}}Y_{\ell}=\ell(\ell+D-2)Y_{\ell}.
\]
The cylinder energy is
\[
  E_\ell
  =
  \sqrt{\ell(\ell+D-2)+\frac{(D-2)^2}{4}}
  =
  \ell+\frac{D-2}{2}.
\]
The \(\ell=0\) mode corresponds to the primary field \(\phi\), whose scaling
dimension is
\[
  \Delta_\phi=\frac{D-2}{2}.
\]
Modes with \(\ell>0\) correspond to derivative descendants of \(\phi\).  This
explicit spectrum realizes the state-operator correspondence in a theory where
all local operators can be built from normal-ordered products of \(\phi\) and
its derivatives, modulo the equation of motion.

\section{The Operator Partition Function of the Free Scalar}

At a fixed point, it is often useful to count linearly independent local
operators by their scaling dimension.  This is a graded trace on the algebraic
local-operator space, not a trace over the Hilbert completion.  Let
\[
  Z_{\rm op}(q)
  =
  \operatorname{Tr}_{\mathcal V_{\rm loc}^{\rm alg}}
  q^{\widehat D_{\rm rad}}
\]
be the formal character of the local-operator space.  In the free scalar
theory, the independent single-letter operators are the symmetric traceless
derivatives
\[
  \partial_{(\mu_1}\cdots\partial_{\mu_\ell)}\phi
  \quad\text{with traces removed},
  \qquad
  \ell=0,1,2,\ldots .
\]
The trace removal is the equation of motion \(\partial^2\phi=0\) and all of
its derivatives.  Counting all symmetric derivatives and subtracting the
traces gives the single-letter character
\begin{equation}
  f_\phi(q)
  =
  q^{\frac{D-2}{2}}
  \frac{1-q^2}{(1-q)^D}.
  \label{eq:free-scalar-letter-partition}
\end{equation}

\begin{proposition}[Free scalar letter character as a short-module character]
\label{prop:free-scalar-letter-short-module-character}
The factor \(1-q^2\) in
\eqref{eq:free-scalar-letter-partition} is the character subtraction of the
null submodule generated by the level-two scalar descendant
\(\widehat P^2\ket{\phi}\), where
\(\widehat P^2=\delta^{\mu\nu}\widehat P_\mu\widehat P_\nu\).  Equivalently,
\[
  f_\phi(q)
  =
  \chi\!\left(M\!\left(\frac{D-2}{2},\mathbf 1\right)\right)
  -
  q^2\,
  \chi\!\left(M\!\left(\frac{D-2}{2},\mathbf 1\right)\right),
\]
where \(M(\Delta,\mathbf 1)\) is the scalar conformal generalized Verma module
defined in Chapter~\ref{ch:volume-iii-primary-operators-finite-transformations}
and \(\mathbf 1\) is the trivial \(SO(D)\) representation.
This is the same null relation as the scalar unitarity-bound shortening
proved in Theorem~\ref{thm:scalar-unitarity-bound-shortening}.
\end{proposition}

\begin{proof}
Before quotienting null states, the scalar generalized Verma module is freely
generated by the commuting translations \(\widehat P_\mu\).  Hence its graded
character is
\[
  \chi(M(\Delta,\mathbf 1))
  =
  q^\Delta\sum_{n=0}^{\infty}
  \dim\operatorname{Sym}^n(\mathbb C^D)\,q^n
  =
  \frac{q^\Delta}{(1-q)^D}.
\]
At \(\Delta_\phi=(D-2)/2\), the level-two trace descendant is null by
Theorem~\ref{thm:scalar-unitarity-bound-shortening}.  It is also a singular
vector.  Indeed, using
\[
  [\widehat K_\nu,\widehat P_\mu]
  =
  2\delta_{\nu\mu}\widehat D_{\rm rad}
  -2\widehat J_{\mu\nu},
  \qquad
  [\widehat J_{\mu\nu},\widehat P_\rho]
  =
  \delta_{\mu\rho}\widehat P_\nu-\delta_{\nu\rho}\widehat P_\mu,
\]
and \(\widehat J_{\mu\nu}\ket{\phi}=0\), one obtains
\begin{align*}
  \widehat K_\nu \widehat P^2\ket{\phi}
  &=
  \left(
    [\widehat K_\nu,\widehat P_\mu]\widehat P_\mu
    +
    \widehat P_\mu[\widehat K_\nu,\widehat P_\mu]
  \right)\ket{\phi}
  \\
  &=
  \left(
    2(\Delta_\phi+1)\widehat P_\nu
    -2(D-1)\widehat P_\nu
    +2\Delta_\phi \widehat P_\nu
  \right)\ket{\phi}
  \\
  &=
  4\left(\Delta_\phi+1-\frac D2\right)\widehat P_\nu\ket{\phi}
  =
  0 .
\end{align*}
Therefore the null vector generates a conformal submodule.  Its translation
descendants have character
\[
  \frac{q^{\Delta_\phi+2}}{(1-q)^D}.
\]

In polynomial notation \(\widehat P_\mu\leftrightarrow p_\mu\), this quotient
is the quotient of \(\mathbb C[p_1,\ldots,p_D]\) by the ideal
\(p^2\mathbb C[p_1,\ldots,p_D]\), where
\(p^2=\delta^{\mu\nu}p_\mu p_\nu\).  At homogeneous degree \(\ell\), the
Fischer decomposition gives
\[
  \operatorname{Sym}^\ell(\mathbb C^D)
  =
  \mathcal H_\ell
  \oplus
  p^2\operatorname{Sym}^{\ell-2}(\mathbb C^D),
\]
where \(\mathcal H_\ell\) is the space of harmonic homogeneous polynomials,
equivalently the symmetric traceless rank-\(\ell\) representation of
\(SO(D)\).  Thus quotienting by the null submodule leaves exactly the
symmetric traceless derivatives of \(\phi\), and the character is
\[
  \frac{q^{\Delta_\phi}}{(1-q)^D}
  -
  \frac{q^{\Delta_\phi+2}}{(1-q)^D}
  =
  q^{\frac{D-2}{2}}\frac{1-q^2}{(1-q)^D}.
\]
\end{proof}

Equivalently,
\[
  f_\phi(q)
  =
  q^{\frac{D-2}{2}}
  \sum_{\ell=0}^\infty d_{\ell,D}q^\ell,
  \qquad
  d_{\ell,D}
  =
  \binom{D+\ell-1}{\ell}
  -
  \binom{D+\ell-3}{\ell-2},
\]
with the second binomial interpreted as zero for \(\ell<2\).  These
\(d_{\ell,D}\) are the dimensions of scalar spherical harmonics of degree
\(\ell\) on \(S^{D-1}\).

Normal-ordered products of the letters form the bosonic symmetric algebra of
single-particle letters.  If
\[
  f_\phi(q)=\sum_\Delta a_\Delta q^\Delta,
\]
then the free scalar operator partition function is
\begin{equation}
  Z_{\rm op}^{\phi}(q)
  =
  \prod_\Delta
  \frac1{(1-q^\Delta)^{a_\Delta}}
  =
  \exp\left(
    \sum_{m=1}^{\infty}\frac1m f_\phi(q^m)
  \right).
  \label{eq:free-scalar-operator-partition}
\end{equation}
For \(D=3\), \(d_{\ell,3}=2\ell+1\), and therefore
\[
  f_\phi(q)
  =
  \sum_{\ell=0}^{\infty}(2\ell+1)q^{\ell+\frac12}
  =
  q^{1/2}\frac{1-q^2}{(1-q)^3}.
\]
This is the same counting obtained from the one-particle spectrum of
\(\widehat H_{\rm cyl}=\sqrt{-\nabla_{S^2}^2+1/4}\) on \(S^2\).

\section{The Free Fermion Operator Partition Function}
\label{sec:free-fermion-cylinder-partition}

Let \(S\) be a complex spinor representation of \(\operatorname{Spin}(D)\) with dimension
\(s_S=\dim S\).  A free spinor primary
\(\psi_\alpha\), \(\alpha=1,\ldots,s_S\), has scaling dimension
\[
  \Delta_\psi=\frac{D-1}{2}.
\]
The flat equation of motion is the Dirac equation
\begin{equation}
  \gamma^\mu{}_{\beta}{}^\alpha\partial_\mu\psi_\alpha=0,
  \label{eq:free-fermion-dirac-equation}
\end{equation}
where \(\gamma^\mu:V\otimes S\to S'\) is Clifford multiplication from the
vector representation \(V\simeq\mathbb C^D\) and the chosen spinor module
\(S\) to the target spinor module \(S'\).  In even \(D\), for a chiral Weyl
field, \(S'\) is the opposite chirality spinor; for a Dirac spinor,
\(S'=S\) after including both chiralities.  The unrefined character only uses
\(\dim S'=\dim S=s_S\).

Before imposing \eqref{eq:free-fermion-dirac-equation}, the translation
descendants form
\[
  S\otimes\mathbb C[p_1,\ldots,p_D],
  \qquad
  p_\mu\leftrightarrow\widehat P_\mu,
\]
with character \(s_S q^{\Delta_\psi}/(1-q)^D\).  The Dirac equation is a
level-one singular vector: the null descendant
\(\gamma^\mu\widehat P_\mu\ket{\psi}\) generates a submodule with character
\(s_S q^{\Delta_\psi+1}/(1-q)^D\).  Therefore the fermion single-letter
character is
\begin{equation}
  f_\psi(q)
  =
  s_S q^{\frac{D-1}{2}}\frac{1-q}{(1-q)^D}
  =
  \frac{s_S q^{\frac{D-1}{2}}}{(1-q)^{D-1}}.
  \label{eq:free-fermion-letter-character}
\end{equation}

\begin{proposition}[Free fermion letter character as a short-module character]
\label{prop:free-fermion-short-module-character}
The character \eqref{eq:free-fermion-letter-character} is the character of
the free spinor conformal module obtained by quotienting the spinor
generalized Verma module by the submodule generated by the Dirac descendant
\(\gamma^\mu\widehat P_\mu\ket{\psi}\).
\end{proposition}

\begin{proof}
The generalized Verma module before shortening is freely generated by the
commuting translations:
\[
  M(\Delta_\psi,S)
  =
  \mathbb C[\widehat P_1,\ldots,\widehat P_D]\otimes S,
  \qquad
  \chi(M(\Delta_\psi,S))
  =
  \frac{s_S q^{\Delta_\psi}}{(1-q)^D}.
\]
At \(\Delta_\psi=(D-1)/2\), the Dirac descendant is primary.  This follows
from the radial commutator
\[
  [\widehat K_\nu,\widehat P_\mu]
  =
  2\delta_{\nu\mu}\widehat D_{\rm rad}
  -2\widehat J_{\mu\nu}
\]
and the spinor action
\[
  \widehat J_{\mu\nu}\ket{\psi}
  =
  \Sigma_{\mu\nu}\ket{\psi},
  \qquad
  \Sigma_{\mu\nu}=\frac14[\gamma_\mu,\gamma_\nu].
\]
Using
\[
  \gamma^\mu\Sigma_{\mu\nu}
  =
  \frac{D-1}{2}\gamma_\nu
\]
with the displayed \(\Sigma_{\mu\nu}\) convention gives
\[
  \widehat K_\nu\gamma^\mu\widehat P_\mu\ket{\psi}
  =
  2\left(\Delta_\psi-\frac{D-1}{2}\right)
  \gamma_\nu\ket{\psi}
  =
  0.
\]
Hence the Dirac descendant generates a conformal submodule.  The submodule is
again freely generated by translations, now starting one level higher and
with spinor multiplicity \(s_S\).  Subtracting its character gives
\eqref{eq:free-fermion-letter-character}.
\end{proof}

If \(f_\psi(q)=\sum_\Delta a_\Delta q^\Delta\), the local-operator algebra
generated by one free fermion field uses the exterior algebra of the odd
letters:
\begin{equation}
  Z_{\rm op}^{\psi}(q)
  =
  \prod_\Delta(1+q^\Delta)^{a_\Delta}
  =
  \exp\left(
    \sum_{m=1}^{\infty}\frac{(-1)^{m+1}}{m}f_\psi(q^m)
  \right).
  \label{eq:free-fermion-operator-partition}
\end{equation}
This is the ordinary positive-counting trace over the fermionic Fock space;
the insertion of \((-1)^F\) would replace \(1+q^\Delta\) by
\(1-q^\Delta\).

In four dimensions, a Weyl fermion has \(s_S=2\) and
\[
  f_{\rm Weyl}(q)=\frac{2q^{3/2}}{(1-q)^3}.
\]
A Dirac fermion, or a Weyl fermion together with its conjugate, has
\[
  f_{\rm Dirac}(q)=\frac{4q^{3/2}}{(1-q)^3}.
\]
The coefficient of \(q^{\ell+3/2}\) for a Weyl fermion is
\((\ell+1)(\ell+2)\), the degeneracy of one chiral tower of spinor harmonics
on \(S^3\).

\section{The Free Maxwell Operator Partition Function}
\label{sec:free-maxwell-cylinder-partition}

The Maxwell field is conformal in four spacetime dimensions.  The local
gauge-invariant letters are the field-strength components \(F_{\mu\nu}\).
Work in Euclidean \(D=4\) with vector representation
\(V\simeq\mathbb C^4\).  The primary \(F_{\mu\nu}\in\wedge^2V\) has
\(\Delta_F=2\) and \(\dim\wedge^2V=6\).  Before imposing the equations, its
translation descendants have character
\[
  \frac{6q^2}{(1-q)^4}.
\]
The Bianchi identity and Maxwell equation are
\begin{equation}
  \partial_{[\lambda}F_{\mu\nu]}=0,
  \qquad
  \partial^\mu F_{\mu\nu}=0.
  \label{eq:maxwell-bianchi-eom-letter}
\end{equation}
At the level of polynomial jets, these are respectively a three-form
constraint and a one-form constraint.  Each has dimension \(4\), so together
they subtract \(8q^3/(1-q)^4\).  The two constraints have one scalar
reducibility identity each:
\[
  \partial_{[\rho}\partial_\lambda F_{\mu\nu]}=0,
  \qquad
  \partial^\nu\partial^\mu F_{\mu\nu}=0.
\]
The first is \(d^2F=0\); the second follows from the antisymmetry of
\(F_{\mu\nu}\).  They restore \(2q^4/(1-q)^4\).  Thus the Maxwell
single-letter character is
\begin{equation}
  f_{\rm Max}(q)
  =
  \frac{q^2(6-8q+2q^2)}{(1-q)^4}
  =
  \frac{6q^2-2q^3}{(1-q)^3}.
  \label{eq:free-maxwell-letter-character}
\end{equation}

Equivalently, decompose
\[
  F=F^++F^-,
  \qquad
  F^+\in(1,0),
  \qquad
  F^-\in(0,1)
\]
under \(\operatorname{Spin}(4)\simeq SU(2)_L\times SU(2)_R\).  Each chiral field-strength
module has character
\[
  f_{F^\pm}(q)
  =
  \frac{q^2(3-4q+q^2)}{(1-q)^4},
  \qquad
  f_{\rm Max}(q)=f_{F^+}(q)+f_{F^-}(q).
\]
This chiral expression makes the conformal shortening transparent: the
level-one divergence descendant is removed, and the level-two scalar
reducibility identity restores the over-subtraction.

If
\[
  f_{\rm Max}(q)=\sum_\Delta b_\Delta q^\Delta,
\]
then the abelian Maxwell operator partition function generated by
normal-ordered products of field-strength letters is the bosonic symmetric
algebra
\begin{equation}
  Z_{\rm op}^{\rm Max}(q)
  =
  \prod_\Delta\frac1{(1-q^\Delta)^{b_\Delta}}
  =
  \exp\left(
    \sum_{m=1}^{\infty}\frac1m f_{\rm Max}(q^m)
  \right).
  \label{eq:free-maxwell-operator-partition}
\end{equation}
Potential variables \(A_\mu\), gauge parameters, and ghosts belong to a
gauge-fixed complex for path-integral determinants.  The character
\eqref{eq:free-maxwell-letter-character} counts the BRST-invariant local
field-strength cohomology in the free abelian theory.

\section{High-Temperature Operator Entropy and the Cardy Limit}
\label{sec:cardy-operator-entropy}

For a CFT satisfying Hypothesis~\ref{hyp:radial-reconstruction-data}, assume
in addition that the spectral projection of \(D_{\rm rad}\) onto
\([0,\Delta]\) has finite trace for every finite \(\Delta\).  The cylinder
thermal trace is then
\begin{equation}
  Z_{\rm cyl}(\beta)
  =
  \operatorname{Tr}_{\mathcal H_{S^{D-1}}}
  \ee^{-\beta D_{\rm rad}}
  =
  \int_{0}^{\infty}\ee^{-\beta E}\,dN(E),
  \qquad
  \beta>0,
  \label{eq:cylinder-thermal-trace}
\end{equation}
where
\[
  N(\Delta)
  =
  \operatorname{Tr}_{\mathcal H_{S^{D-1}}}
  \mathbf 1_{[0,\Delta]}(D_{\rm rad}).
\]
If the spectrum is pure point with finite degeneracies, this integral is
\(\sum_\Delta d(\Delta)\ee^{-\beta\Delta}\), where
\(d(\Delta)=\dim\ker(D_{\rm rad}-\Delta)\).  The operator partition function
\(Z_{\rm op}(q)\) is this same trace with \(q=\ee^{-\beta}\) when the local
operator basis is complete and the trace converges.  The large-\(\Delta\)
density should therefore be understood as the integrated spectral counting
function \(N(\Delta)\), or as a smoothed density obtained from it, rather than
as an everywhere-defined ordinary function.

Assume that the high-temperature cylinder free energy has the leading form
\begin{equation}
  \log Z_{\rm cyl}(\beta)
  =
  A\,\beta^{-(D-1)}+o(\beta^{-(D-1)}),
  \qquad
  A>0 .
  \label{eq:cylinder-high-temp-free-energy}
\end{equation}
This hypothesis is a thermodynamic input about the CFT on
\(S^{D-1}\): in \(D>2\) the constant \(A\) is not fixed by the stress-tensor
two-point coefficient alone.  It is the finite-volume version of the flat
thermal pressure coefficient.  The precise positivity input is the following
Hardy--Littlewood Tauberian implication: if \(dN\) is a positive spectral
measure on \([0,\infty)\) and
\(\log\int_0^\infty\ee^{-\beta E}dN(E)=A\beta^{-\alpha}+o(\beta^{-\alpha})\)
with \(A,\alpha>0\), then
\[
  \log N(E)
  =
  (\alpha+1)A^{1/(\alpha+1)}
  \left(\frac{E}{\alpha}\right)^{\alpha/(\alpha+1)}
  +o(E^{\alpha/(\alpha+1)}).
\]
Applying this with \(\alpha=D-1\) gives the leading microcanonical entropy
\begin{equation}
  \log N(\Delta)
  =
  D\,A^{1/D}
  \left(\frac{\Delta}{D-1}\right)^{\frac{D-1}{D}}
  +o(\Delta^{\frac{D-1}{D}}).
  \label{eq:general-cft-entropy-dimension-asymptotic}
\end{equation}
The saddle calculation is explicit.  Extremize
\[
  \Phi(\beta)=\beta\Delta+A\beta^{-(D-1)}.
\]
The stationary point satisfies
\[
  \Delta=(D-1)A\beta_\ast^{-D},
  \qquad
  \beta_\ast=\left(\frac{(D-1)A}{\Delta}\right)^{1/D}.
\]
Substitution gives \eqref{eq:general-cft-entropy-dimension-asymptotic}.  Thus
the power \(\Delta^{(D-1)/D}\) is fixed by extensivity of high-temperature
thermal physics in \(D-1\) spatial dimensions, while the coefficient records
the specific CFT thermal free energy.

In \(D=2\) the torus supplies a stronger statement because modular invariance
relates the high-temperature trace to the low-temperature vacuum.  Let the
spatial circle have circumference \(2\pi\), and let
\[
  H_{\rm cyl}=L_0+\bar L_0-\frac{c_L+c_R}{24},
  \qquad
  \Delta=h+\bar h,
  \qquad
  J=h-\bar h .
\]
For a compact unitary two-dimensional CFT with discrete finite-multiplicity
spectrum, unique vacuum, and modular-invariant torus partition function, the
thermal trace
\[
  Z(\beta)
  =
  \operatorname{Tr}
  \exp\left[-\beta\left(L_0+\bar L_0-\frac{c_L+c_R}{24}\right)\right]
\]
obeys \(Z(\beta)=Z(4\pi^2/\beta)\) for the rectangular torus.  As
\(\beta\to0^+\), the transformed trace is dominated by the vacuum, and
\begin{equation}
  \log Z(\beta)
  =
  \frac{\pi^2(c_L+c_R)}{6\beta}
  +o(\beta^{-1}).
  \label{eq:modular-high-temp-cardy}
\end{equation}
If \(c_L=c_R=c\), then \(A=\pi^2c/3\) in
\eqref{eq:cylinder-high-temp-free-energy}.  Equation
\eqref{eq:general-cft-entropy-dimension-asymptotic} gives
\begin{equation}
  \log N(\Delta)
  =
  2\pi\sqrt{\frac{c\,\Delta}{3}}
  +o(\sqrt{\Delta}).
  \label{eq:cardy-total-dimension-asymptotic}
\end{equation}
This is the Cardy entropy-dimension relation for total scaling dimension.
Keeping left and right energies separately gives the spin-resolved form
\begin{equation}
  \log \rho(h,\bar h)
  =
  2\pi\sqrt{\frac{c_L h}{6}}
  +
  2\pi\sqrt{\frac{c_R\bar h}{6}}
  +o(\sqrt h+\sqrt{\bar h}),
  \label{eq:cardy-spin-resolved}
\end{equation}
for \(h,\bar h\to\infty\) with \(h/\bar h\) remaining in a compact subinterval
of \((0,\infty)\).  If the lowest left or right conformal weight is not the
vacuum value, the same derivation replaces \(c_L,c_R\) by the corresponding
effective central charges \(c_L-24h_{\min}\) and
\(c_R-24\bar h_{\min}\).
