\chapter{Light-Ray Operators and Energy Correlators}
\label{ch:cft-light-ray-energy-correlators}

\ConventionsMixedCFT

Energy correlators provide a common observable language for conformal field
theory and high-energy gauge theory.  The primitive object is a detector at
future null infinity, represented in local QFT by an asymptotic flux of the
stress tensor.  The same object can also be described, after conformal
compactification, as a null-integrated light-ray operator.

\section{Detector States and Energy Flux}

Let the Lorentzian CFT live on \(\mathbb M^D\), \(D\geq3\), with Hilbert space
\(\Hilb\), vacuum vector \(\vac\), translation generators \(P^\mu\), and a
conserved symmetric stress tensor \(T^{\mu\nu}\).  A normalizable collider
state is obtained by smearing a local operator with timelike momentum support.
For a local operator \(\mathcal O_A\), where \(A\) denotes all tensor and
internal indices, choose a smooth compactly supported wavepacket
\(h(q)\) on the open future cone and a polarization tensor \(\varepsilon^A\),
and set
\[
  \ket{\Psi_{\mathcal O,h,\varepsilon}}
  =
  \int_{V_+}\dd\mu(q)\,
  h(q)\,\varepsilon^A\,\widetilde{\mathcal O}_A(q)\ket{\vac}.
\]
Here \(\dd\mu(q)\) is a positive measure on the chosen momentum slice.  The
Fourier-transformed operator is an operator-valued distribution applied to the
test function \(h\).  Sharp momentum states used in collider formulas are
distributional limits of these wavepackets.

\begin{hypothesis}[CFT energy detector]
\label{hyp:cft-energy-detector}
For every real \(f\in C^\infty(S^{D-2})\), the limit
\begin{equation}
  \mathcal E(f)
  =
  \mathrm{w}\!-\!\lim_{r\to\infty}
  r^{D-2}
  \int_{-\infty}^{\infty}\dd u
  \int_{S^{D-2}}\dd\Omega(\mathbf n)\,
  f(\mathbf n)\,
  n_i T^{0i}(u+r,r\mathbf n)
  \label{eq:cft-energy-detector}
\end{equation}
exists as a quadratic form on the collider domain generated by such states.
If \(f\geq0\), the resulting quadratic form is positive.  The constant
function gives the total outgoing energy,
\[
  \mathcal E(1)=P^0
\]
on the same domain.
\end{hypothesis}

The distributional detector density is denoted by
\[
  \mathcal E(f)
  =
  \int_{S^{D-2}}\dd\Omega(\mathbf n)\,
  f(\mathbf n)\mathcal E(\mathbf n).
\]
For \(k\) test functions \(f_1,\ldots,f_k\), the \(k\)-point energy correlator
in a normalized state \(\Psi\) is
\begin{equation}
  \mathcal G_k(f_1,\ldots,f_k;\Psi)
  =
  \bra{\Psi}\mathcal E(f_1)\cdots\mathcal E(f_k)\ket{\Psi}.
  \label{eq:cft-energy-correlator}
\end{equation}
The product is defined first for separated angular supports and then extended,
when possible, as a distribution on \((S^{D-2})^k\).  Coincident detector
directions carry contact terms determined by the distributional product of
light-ray operators.

\subsection{Finite Angular Partitions and Positivity}
\label{subsec:cft-detector-finite-partitions-positivity}

Before multiplying detector operators, it is useful to extract the
one-detector positivity consequence of Hypothesis
\ref{hyp:cft-energy-detector}.  Let \(\Psi\) be a finite-energy collider
state in the detector domain and define the state functional
\[
  \ell_\Psi(f)
  =
  \bra{\Psi}\mathcal E(f)\ket{\Psi},
  \qquad f\in C^\infty(S^{D-2}),\quad f\ {\rm real}.
\]
If \(f\geq0\), then \(\ell_\Psi(f)\geq0\).  Moreover
\(-\|f\|_\infty\leq f\leq\|f\|_\infty\) implies, by positivity and
\(\mathcal E(1)=P^0\),
\begin{equation}
  |\ell_\Psi(f)|
  \leq
  \|f\|_\infty\,\bra{\Psi}P^0\ket{\Psi}.
  \label{eq:cft-detector-riesz-bound}
\end{equation}
Thus \(\ell_\Psi\) extends uniquely from smooth tests to a positive bounded
linear functional on \(C(S^{D-2})\).  By the Riesz representation theorem
there is a finite positive regular Borel measure \(\mu_\Psi\) on the detector
sphere such that
\begin{equation}
  \bra{\Psi}\mathcal E(f)\ket{\Psi}
  =
  \int_{S^{D-2}} f(\mathbf n)\,\dd\mu_\Psi(\mathbf n),
  \qquad
  \mu_\Psi(S^{D-2})=\bra{\Psi}P^0\ket{\Psi}.
  \label{eq:cft-statewise-detector-measure}
\end{equation}
This is a statewise measure statement.  Promoting it to an
operator-valued measure requires compatible polarized matrix elements and a
common domain; those domain questions are additional parts of the
detector-existence hypothesis beyond the scalar Riesz representation step.

For a finite angular partition
\(\mathcal P=\{B_1,\ldots,B_m\}\) of the sphere by Borel sets, the numbers
\[
  \mu_a(\Psi)=\mu_\Psi(B_a)\geq0,\qquad
  \sum_{a=1}^m\mu_a(\Psi)=\bra{\Psi}P^0\ket{\Psi}
\]
are the finite-resolution calorimeter readout in the state \(\Psi\).  If
\(v,w\in\mathbb C^m\) are bin weights, the statewise one-detector pairing
\[
  (v,w)_\Psi
  =
  \sum_{a=1}^m \mu_a(\Psi)\,\overline{v_a}w_a
\]
is positive semidefinite and obeys
\begin{equation}
  |(v,w)_\Psi|^2
  \leq
  (v,v)_\Psi\,(w,w)_\Psi .
  \label{eq:cft-detector-bin-cauchy-schwarz}
\end{equation}
These inequalities are consequences of positivity of a single detector
measure.  Positivity statements for products
\(\mathcal E(f_1)\cdots\mathcal E(f_k)\) require the diagonal-contact
extension discussed next.

\paragraph{Finite-resolution Lipschitz estimates.}
The finite partition readout gives the operational topology in which an ideal
detector is approached by finite angular bins.  Let \(d\) be the round metric on
\(S^{D-2}\).  For a
finite Borel partition
\[
  \mathcal P=\{B_1,\ldots,B_m\},
  \qquad
  \operatorname{diam}(\mathcal P)
  =
  \max_a\sup_{\mathbf n,\mathbf n'\in B_a}d(\mathbf n,\mathbf n'),
\]
choose a representative point \(\mathbf n_a\in B_a\) for each nonempty bin and
replace the statewise detector measure by the binned measure
\[
  \mu_{\Psi,\mathcal P}
  =
  \sum_a\mu_\Psi(B_a)\,\delta_{\mathbf n_a}.
\]
If \(f:S^{D-2}\to\mathbb R\) is \(L_f\)-Lipschitz, then
\begin{equation}
  \left|
  \int f\,\dd\mu_\Psi
  -
  \int f\,\dd\mu_{\Psi,\mathcal P}
  \right|
  \leq
  L_f\,\operatorname{diam}(\mathcal P)\,\mu_\Psi(S^{D-2}).
  \label{eq:cft-detector-partition-lipschitz-estimate}
\end{equation}
Indeed, on each bin,
\[
  |f(\mathbf n)-f(\mathbf n_a)|
  \leq
  L_f\,\operatorname{diam}(\mathcal P),
\]
and summing the resulting measure estimate over bins gives
\eqref{eq:cft-detector-partition-lipschitz-estimate}.  Thus convergence of
finite bin masses along partitions with
\(\operatorname{diam}(\mathcal P)\to0\), together with the uniform total
energy bound \(\mu_\Psi(S^{D-2})=\langle P^0\rangle_\Psi\), implies weak
convergence of the statewise detector measures against continuous tests by
uniform approximation.

The same estimate controls finite-resolution \(k\)-detector products after a
positive joint detector measure has been constructed.  Let
\(\mu_\Psi^{(k)}\) be such a measure on \((S^{D-2})^k\), let
\(\mu_{\Psi,\mathcal P}^{(k)}\) be the pushforward to the representative
points, and let \(F\) be \(L_F\)-Lipschitz with respect to the product metric
\[
  d_k((\mathbf n_1,\ldots,\mathbf n_k),
      (\mathbf n'_1,\ldots,\mathbf n'_k))
  =
  \sum_{r=1}^{k}d(\mathbf n_r,\mathbf n'_r).
\]
Then
\begin{equation}
  \left|
  \int F\,\dd\mu_\Psi^{(k)}
  -
  \int F\,\dd\mu_{\Psi,\mathcal P}^{(k)}
  \right|
  \leq
  k\,L_F\,\operatorname{diam}(\mathcal P)\,
  \mu_\Psi^{(k)}((S^{D-2})^k).
  \label{eq:cft-detector-product-partition-lipschitz-estimate}
\end{equation}
For the product measure of a finite-energy state, the total mass is
\(\langle(P^0)^k\rangle_\Psi\) when the diagonal-contact extension has been
included.  Equations
\eqref{eq:cft-detector-partition-lipschitz-estimate} and
\eqref{eq:cft-detector-product-partition-lipschitz-estimate} are therefore
the finite-resolution Lipschitz estimates behind the statement that energy
correlators are detector observables on the celestial sphere.  The
construction of detector products remains the separate domain question; once a
product measure is available, these estimates identify the finite calorimeter
data it supplies.

\subsection{Angular Moment Data and Measure Reconstruction}
\label{subsec:cft-detector-moment-reconstruction}

The statewise measure \(\mu_\Psi\) turns angular detector data into a compact
moment problem.  Let \(K=S^{D-2}\), and let
\(\mathcal A_{\rm ang}\subset C(K;\mathbb R)\) be a unital real subalgebra
that separates points and is uniformly dense in \(C(K;\mathbb R)\).  A
concrete choice is the algebra generated by the restrictions of the ambient
coordinate functions \(n^i\) to \(S^{D-2}\); equivalently, one may use any
real spherical-harmonic basis closed under finite multiplication.  The
angular moments of a finite detector measure are the numbers
\begin{equation}
  m_\Psi(a)
  =
  \int_K a(\mathbf n)\,\dd\mu_\Psi(\mathbf n),
  \qquad a\in\mathcal A_{\rm ang}.
  \label{eq:cft-detector-angular-moments}
\end{equation}
For \(a=1\) this moment is the total energy
\(\bra{\Psi}P^0\ket{\Psi}\).

\paragraph{Compact angular moment reconstruction.}
\label{par:cft-detector-compact-moment-reconstruction}
Let \(M:\mathcal A_{\rm ang}\to\mathbb R\) be linear and assume
\[
  M(a)\geq0
  \quad\text{whenever}\quad
  a(\mathbf n)\geq0\ \text{for all }\mathbf n\in K,
  \qquad
  M(1)=E<\infty .
\]
Then there is a unique finite positive regular Borel measure \(\mu_M\) on
\(K\) satisfying
\[
  M(a)=\int_K a\,\dd\mu_M,
  \qquad a\in\mathcal A_{\rm ang}.
\]
Consequently the complete angular moment functional
\(a\mapsto m_\Psi(a)\) determines the statewise detector measure
\(\mu_\Psi\) uniquely.

\emph{Argument.}
For real \(a\in\mathcal A_{\rm ang}\), the inequalities
\(-\|a\|_\infty\leq a\leq\|a\|_\infty\) on \(K\) imply
\[
  0\leq M(\|a\|_\infty 1\pm a)
  =
  \|a\|_\infty E\pm M(a).
\]
Hence \(|M(a)|\leq E\|a\|_\infty\).  Thus \(M\) is uniformly bounded on the
dense subalgebra \(\mathcal A_{\rm ang}\) and extends uniquely by continuity
to a bounded positive linear functional \(\overline M\) on
\(C(K;\mathbb R)\).  Positivity of the extension follows by approximating any
nonnegative \(f\in C(K;\mathbb R)\) uniformly by \(a_j\in\mathcal A_{\rm ang}\)
and applying the bound to \(a_j-f\) together with the positivity assumption
to the functions \(a_j+\epsilon_j 1\), where
\(\epsilon_j=\|a_j-f\|_\infty\).

The Riesz--Markov representation theorem gives a finite positive regular
Borel measure \(\mu_M\) with
\(\overline M(f)=\int_K f\,\dd\mu_M\) for all continuous real \(f\).  The
restriction to \(\mathcal A_{\rm ang}\) gives the displayed moment identities.
If two finite positive measures have the same moments on
\(\mathcal A_{\rm ang}\), their integrals agree on a uniformly dense subset
of \(C(K)\).  The bound by total mass extends this equality to all continuous
functions, and the uniqueness clause in the Riesz--Markov theorem makes the
two measures equal.

This criterion records the information content of complete angular detector
moments.  A finite list of moments is a finite-resolution projection of
\(\mu_\Psi\).  It gives positivity constraints and linear coordinates for a
specified detector resolution, while many distinct positive measures may
share the same truncated list.  For example on any detector great circle, the
measures \(\frac12\delta_{\mathbf n}+\frac12\delta_{-\mathbf n}\) and
\(\delta_{\mathbf n_\perp}\), with
\(\mathbf n_\perp\cdot\mathbf n=0\), have the same total energy and the same
dipole moment but different quadrupole moments.  The complete moment
functional separates them because \(\mathcal A_{\rm ang}\) separates points.

The same compact moment logic applies to detector products after the product
has been extended to a finite positive joint detector measure
\(\mu_\Psi^{(k)}\) on \(K^k\).  Then moments of
\(\mu_\Psi^{(k)}\) against the algebraic tensor products of
\(\mathcal A_{\rm ang}\) determine that \(k\)-detector measure uniquely.  The
substantive QFT problem is the construction of the product measure and its
diagonal-contact coordinates; once that finite measure is supplied, compact
moment reconstruction is ordinary functional analysis.  The QCD
calorimetric-coordinate density theorem,
Theorem~\ref{thm:energy-correlator-polynomial-density}, is the corresponding
statement on the compact ball of finite positive energy measures.

\subsection{Detector Products and Diagonal Contacts}
\label{subsec:cft-detector-products-contacts}

The contact terms in detector products have a concrete measure-theoretic
origin.  A finite-resolution calorimetric configuration \(X\) with detector
directions \(\mathbf n_a\in S^{D-2}\) and positive energies \(E_a>0\)
determines the positive finite measure
\begin{equation}
  \varepsilon_X
  =
  \sum_{a=1}^N E_a\,\delta_{\mathbf n_a}.
  \label{eq:finite-detector-energy-measure}
\end{equation}
The corresponding detector functional is
\[
  E_X(f)=\int f\,\dd\varepsilon_X
  =
  \sum_a E_a f(\mathbf n_a).
\]
For two detector tests \(f,g\), the product decomposes exactly as
\begin{align}
  E_X(f)E_X(g)
  &=
  \sum_{a,b}E_aE_b f(\mathbf n_a)g(\mathbf n_b)
  \notag\\
  &=
  \sum_{a\neq b}E_aE_b f(\mathbf n_a)g(\mathbf n_b)
  +
  \sum_a E_a^2 f(\mathbf n_a)g(\mathbf n_a).
  \label{eq:finite-detector-diagonal-decomposition}
\end{align}
The first term is the restriction of the product measure
\(\varepsilon_X\otimes\varepsilon_X\) to the complement of the diagonal in
\(S^{D-2}\times S^{D-2}\).  The second term is supported on the diagonal
\(\mathbf n_1=\mathbf n_2\).  Thus a separated-angle two-detector
distribution and a full two-detector distribution differ by a specified
diagonal distribution; deleting or retaining that diagonal part is a choice of
observable coordinate, not a change of the underlying detector.

For \(k\) detector insertions, the same bookkeeping is indexed by partitions.
Let \(\mathsf P_k\) be the set of set partitions of \(\{1,\ldots,k\}\).  If
\(\pi\in\mathsf P_k\) and \(B\in\pi\) is a block, write \(a_B\) for the
detector atom assigned to that block.  Grouping equal labels in the product
\(\prod_i\sum_{a_i}E_{a_i}f_i(\mathbf n_{a_i})\) gives
\begin{equation}
  \prod_{i=1}^k E_X(f_i)
  =
  \sum_{\pi\in\mathsf P_k}
  \sum_{\substack{a_B,\ B\in\pi\\ a_B\neq a_{B'}\ {\rm for}\ B\neq B'}}
  \prod_{B\in\pi}
  E_{a_B}^{|B|}
  \prod_{i\in B} f_i(\mathbf n_{a_B}).
  \label{eq:finite-detector-partition-decomposition}
\end{equation}
The discrete partition into singletons is the fully separated detector
support.  Every coarser partition is a diagonal contact stratum in
\((S^{D-2})^k\).  Formula
\eqref{eq:finite-detector-partition-decomposition} is the finite positive
measure model for the contact-term structure of
\eqref{eq:cft-energy-correlator}.

In the operator language of the CFT, Hypothesis
\ref{hyp:cft-energy-detector} supplies the smeared detector as a quadratic
form.  Products with mutually disjoint angular supports are defined by the
off-diagonal part of the operator-valued measure.  Extending the product to
overlapping supports requires a distributional extension across the
diagonals, and the extension must preserve the stress-tensor Ward identity
\begin{equation}
  \mathcal G_2(1,1;\Psi)
  =
  \bra{\Psi}(P^0)^2\ket{\Psi}.
  \label{eq:detector-product-total-energy-ward}
\end{equation}
For a state with sharp total energy \(Q\), the normalized two-detector
correlator therefore has unit zeroth moment when the diagonal contact
coordinate is included.  A convention that studies only separated detector
directions must display the removed diagonal contribution separately.  This is
the CFT counterpart of the endpoint-contact bookkeeping for the QCD
energy-energy correlator in
Section~\ref{sec:energy-correlators-calorimetric-coordinates}.
The constant test fixes only the total diagonal mass.  Finite angular bins
require the full contact measure along the diagonal: moving the same contact
mass from one bin to another preserves
\(\mathcal G_2(1,1;\Psi)\), but it changes the product
\(\mathcal G_2(\mathbf 1_B,\mathbf 1_B;\Psi)\) for a localized calorimeter
bin \(B\).  Thus the contact coordinate is a measure-valued part of the
detector distribution, with one coordinate for each diagonal angular
resolution cell.

\section{Relation to Null-Integrated Operators}

Let \(n^\mu=(1,\mathbf n)\) be a future null vector.  For a point
\(x_\perp\) in a spacelike plane transverse to \(n^\mu\), the averaged null
energy operator on the null line through \(x_\perp\) is the quadratic form
\begin{equation}
  \mathcal A_n(x_\perp)
  =
  \int_{-\infty}^{\infty}\dd\lambda\,
  T_{\mu\nu}(x_\perp+\lambda n)n^\mu n^\nu .
  \label{eq:anec-light-ray-operator}
\end{equation}
The conformal map from a null line to future null infinity identifies
\(\mathcal E(\mathbf n)\), up to the fixed conformal Jacobian determined by
the chosen compactification, with the light-ray transform
\eqref{eq:anec-light-ray-operator} of the stress tensor.  Thus the detector
operator is a nonlocal operator obtained by integrating a local field along a
complete null generator.

This formulation separates three pieces of data:
\[
  \begin{gathered}
  \text{local stress tensor}
  \longrightarrow
  \text{null-integrated light-ray operator}
  \\
  \longrightarrow
  \text{calorimetric distribution on }S^{D-2}.
  \end{gathered}
\]
The first belongs to the local operator algebra.  The second requires control
of a null integral of an operator-valued distribution.  The third is a
detector observable in states with finite total energy.

For \(f\geq0\), positivity of \(\mathcal E(f)\) is the detector version of the
averaged null energy theorem stated below.  Hypothesis
\ref{hyp:cft-energy-detector} packages two analytic inputs: the existence of
the null-infinity detector as a quadratic form and the positivity supplied by
the theorem for nonnegative angular smearings.

\section{The Light Transform of a Local Primary}
\label{sec:cft-light-transform-local-primary}

The stress-tensor detector is one member of a broader class of light-ray
operators.  We spell out the conformal bookkeeping because the same
representation-theoretic data are used later in the light-ray OPE, and because
it prevents a common ambiguity: a light-ray operator is not an ordinary local
operator placed at a point of spacetime.  It is an integral transform of a
local operator-valued distribution, with its own domain and growth
hypotheses.

Use the standard embedding-space notation for a symmetric traceless primary
\(\mathcal O_{\Delta,J}\).  A physical point is represented by a null ray
\(P\in\mathbb R^{D,2}\), and tensor indices are packaged by a null auxiliary
vector \(Z\) satisfying
\[
  P^2=Z^2=P\cdot Z=0,\qquad Z\sim Z+\beta P .
\]
The encoded primary is homogeneous as
\begin{equation}
  \mathcal O_{\Delta,J}(\lambda P,\alpha Z)
  =
  \lambda^{-\Delta}\alpha^J\,
  \mathcal O_{\Delta,J}(P,Z),
  \qquad \lambda,\alpha>0 .
  \label{eq:embedding-primary-homogeneity}
\end{equation}
This is a compact way to say that \(\Delta\) is the scaling dimension and
\(J\) is the spin.  Gauge invariance under \(Z\to Z+\beta P\) encodes
tracelessness and transversality.

For a primary satisfying the Lorentzian growth hypotheses needed to make the
integral a distribution, its light transform is
\begin{equation}
  \mathbb L[\mathcal O_{\Delta,J}](P,Z)
  =
  \int_{-\infty}^{\infty}\dd\alpha\,
  \mathcal O_{\Delta,J}(Z-\alpha P,-P).
  \label{eq:cft-light-transform-embedding}
\end{equation}
The point \(Z-\alpha P\) runs along a null generator, while the new auxiliary
polarization is \(-P\).  The formula is first a pairing with test functions
in the transverse and ray variables; convergence or analytic continuation is
part of the light-transform datum, not an automatic consequence of local
operator covariance.

The homogeneity of \(\mathbb L[\mathcal O_{\Delta,J}]\) follows directly from
\eqref{eq:embedding-primary-homogeneity}.  Scaling \(P\) and changing
variables \(\alpha'=\lambda\alpha\) gives
\begin{equation}
  \mathbb L[\mathcal O](\lambda P,Z)
  =
  \int\frac{\dd\alpha'}{\lambda}\,
  \mathcal O(Z-\alpha'P,-\lambda P)
  =
  \lambda^{J-1}\mathbb L[\mathcal O](P,Z).
  \label{eq:light-transform-p-homogeneity}
\end{equation}
Scaling \(Z\) and changing variables \(\alpha=\beta\alpha'\) gives
\begin{equation}
  \mathbb L[\mathcal O](P,\beta Z)
  =
  \int\beta\,\dd\alpha'\,
  \mathcal O(\beta(Z-\alpha'P),-P)
  =
  \beta^{1-\Delta}\mathbb L[\mathcal O](P,Z).
  \label{eq:light-transform-z-homogeneity}
\end{equation}
Thus the light transform has conformal weights
\begin{equation}
  (\Delta,J)
  \longmapsto
  (1-J,\,1-\Delta),
  \label{eq:light-transform-weight-map}
\end{equation}
where the second entry is understood as the analytically continued spin
coordinate of the nonlocal light-ray representation.  Equations
\eqref{eq:light-transform-p-homogeneity}--\eqref{eq:light-transform-weight-map}
are representation-theoretic bookkeeping, not an existence theorem.  The
existence theorem requires the Lorentzian analyticity and growth assumptions
appearing in Quoted Theorem~\ref{qthm:light-ray-ope-boundary}.
The signs in \eqref{eq:light-transform-weight-map} are tied to the conformal
charge convention of Lemma~\ref{lem:euclidean-conformal-charge-algebra}, not
to an independent convention for detector operators.  In embedding space the
conformal generators act as linear first-order vector fields on the pair
\((P,Z)\).  Acting on
\(\mathcal O_{\Delta,J}(Z-\alpha P,-P)\) therefore gives the same generator
acting on the transformed line, plus a total derivative in the line coordinate
\(\alpha\).  Under the growth hypotheses that make the light transform a
distribution, that total derivative has no endpoint contribution after
pairing with line test functions.  The transform consequently intertwines the
complexified conformal algebra, while its projective weights change only by
the two Jacobian factors displayed above: \(P\)-rescaling contributes the
\(\dd\alpha/\lambda\) factor and \(Z\)-rescaling contributes the
\(\beta\,\dd\alpha'\) factor.  The radial real form used for descendant Gram
matrices is a separate convention fixed in
\eqref{eq:lorentzian-to-radial-generator-map}--\eqref{eq:radial-pk-sign-from-lorentzian};
it is obtained by the displayed change of real form before Gram matrices are
computed.

For the stress tensor, \((\Delta,J)=(D,2)\).  The light transform therefore
has formal embedding weights \((-1,1-D)\).  After conformal compactification
and restriction to future null infinity, this object is represented by the
positive energy detector density \(\mathcal E(\mathbf n)\), a measure-valued
operator on \(S^{D-2}\) rather than a local primary on Minkowski space.  This
is the reason the detector normalization is fixed by
\(\mathcal E(1)=P^0\) and by the ANEC theorem, not by treating
\(\mathcal E(\mathbf n)\) as an ordinary local operator.

\section{ANEC and Conformal Collider Bounds}
\label{sec:anec-conformal-collider-bounds}

The averaged null energy condition used in conformal collider physics is a
quadratic-form statement about a null light transform of the stress tensor.
It is stronger than positivity of the Hamiltonian, because it localizes the
energy measurement on one complete null generator rather than integrating
over a spacelike Cauchy surface.  The statement requires a domain on which the
null integral of the stress-tensor distribution exists.

Let \(N_n=n^\perp/\mathbb R n\) be the \((D-2)\)-dimensional transverse
quotient to the null direction \(n^\mu\).  For a nonnegative test function
\(\varphi\in C_c^\infty(N_n)\), define the smeared null-energy quadratic form
\begin{equation}
  \mathcal A_n(\varphi)
  =
  \int_{N_n}\dd^{D-2}x_\perp\,\varphi(x_\perp)
  \int_{-\infty}^{\infty}\dd\lambda\,
  T_{\mu\nu}(x_\perp+\lambda n)n^\mu n^\nu .
  \label{eq:anec-smeared-transverse}
\end{equation}
The transverse smearing is part of the definition.  The unsmeared
\(\mathcal A_n(x_\perp)\) in \eqref{eq:anec-light-ray-operator} is the
corresponding distributional kernel.

\begin{quotedtheorem}[Route-scoped ANEC theorem boundary for collider states]
\label{qthm:anec-cft}
Consider a unitary Lorentzian CFT on \(\mathbb M^D\), \(D\geq3\), with a
unique translation-invariant vacuum \(\vac\), spectrum condition, local
stress tensor \(T_{\mu\nu}\) as a Wightman distribution.  Suppose that a
specified ANEC proof route supplies a state class
\(\mathcal D_{\rm ANEC}\), a common quadratic-form domain for the transverse
smearings in \eqref{eq:anec-smeared-transverse}, and regulator-independent
meaning for the complete null integral.  In the modular route this means the
Araki-relative-entropy, null-cut shape-derivative, collar-removal, and common
domain hypotheses used in the deformed-half-space proof.  In the causality
route this means the interacting \(D>2\) Lorentzian-CFT hypotheses, absence of
leading higher-spin degeneracy in the light-cone OPE channel, the
stress-tensor projection, convergence or controlled asymptotic use of the OPE,
and the required analyticity and order-of-limits assumptions.  Then, for every
future null \(n^\mu\), every nonnegative
\(\varphi\in C_c^\infty(N_n)\), and every
\(\Psi\in\mathcal D_{\rm ANEC}\),
\begin{equation}
  \bra{\Psi}\mathcal A_n(\varphi)\ket{\Psi}\geq0 .
  \label{eq:anec-positivity}
\end{equation}
\end{quotedtheorem}

The theorem is used here as a route-scoped external input, not as a consequence
of an unspecified analyticity assumption.\footnote{For the
modular-Hamiltonian proof, see
T.~Faulkner, R.~G.~Leigh, O.~Parrikar, and H.~Wang, ``Modular Hamiltonians
for Deformed Half-Spaces and the Averaged Null Energy Condition,''
JHEP \textbf{09} (2016) 038, arXiv:1605.08072.  For the Lorentzian
causality proof in CFT, see T.~Hartman, S.~Kundu, and A.~Tajdini,
``Averaged null energy condition from causality,'' JHEP \textbf{07} (2017)
066, arXiv:1610.05308.}
Faulkner--Leigh--Parrikar--Wang use regulated modular Hamiltonians and
relative entropy to obtain the null-energy sign; Hartman--Kundu--Tajdini use
the light-cone OPE and causality sum rule under the interacting-theory
assumptions just listed.  The two proof mechanisms have the same collider
output, but they are not interchangeable theorem statements.

Sharp local algebras are type III.  Thus the continuum notation below is
algebraic: for a null-cut region \(R_f\) write
\(\mathfrak M_f=\mathfrak A(R_f)\), let
\(\omega_0\) be the vacuum state and \(\omega_\Psi\) the vector state of
\(\Psi\), and use Araki relative entropy
\(S_{\mathfrak M_f}(\omega_\Psi\Vert\omega_0)\).  The modular group is the
Tomita--Takesaki modular group of \((\mathfrak M_f,\vac)\), with full modular
generator denoted \(\widehat K_f\).  There is no sharp-region reduced density
matrix, trace over \(R_f\), or separately defined local von Neumann entropy in
this notation.

We spell out the null-cut sign convention because conformal-collider
positivity depends on it.  Choose light-cone coordinates
\[
  x^\pm=x^0\pm x^1,\qquad
  y=(x^2,\ldots,x^{D-1}),\qquad
  \dd s^2=-\dd x^+\dd x^-+\dd y^2,
\]
and take \(n=\partial_+\).  For a smooth cut \(x^+=f(y)\) of the null plane
\(x^-=0\), let \(R_f\) be the domain of dependence of the portion
\(x^-\!=0,\ x^+\geq f(y)\).

\begin{quotedtheorem}[Null-cut modular-flow theorem boundary]
Under the CFT null-cone/OPE hypotheses, or under the algebraic
half-sided-modular-inclusion hypotheses for the corresponding null-cut
algebras, the vacuum full modular generator has the following horizon
quadratic-form representation after transverse smearing:
\begin{equation}
  \widehat K_f^{\rm hor}
  =
  2\pi
  \int\dd^{D-2}y
  \int_{-\infty}^{\infty} \dd x^+\,
  \bigl(x^+-f(y)\bigr)T_{++}(x^+,x^-=0,y).
  \label{eq:null-cut-modular-hamiltonian}
\end{equation}
The one-sided future-horizon and past-horizon pieces are obtained by
restricting the \(x^+\)-integral to \(x^+\geq f(y)\) and
\(x^+\leq f(y)\), respectively.  The full generator
\(\widehat K_f\) is the algebraic modular generator; it is not
the logarithm of a sharp local density matrix.
\end{quotedtheorem}

This theorem boundary is the result of Casini--Teste--Torroba's null-plane
analysis and its algebraic half-sided-modular-inclusion version.\footnote{See
H.~Casini, E.~Teste, and G.~Torroba, ``Modular Hamiltonians on the null plane
and the Markov property of the vacuum state,'' J. Phys. A \textbf{50} (2017)
364001, arXiv:1703.10656.}  We use it here to fix the sign of a null
displacement; it is not silently substituted for the ANEC theorem above.
For \(f_s(y)=s\varphi(y)\), differentiation of the full horizon generator at
\(s=0\) gives
\begin{equation}
  \frac{\dd}{\dd s}\,
  \Delta_\Psi\langle \widehat K_{f_s}^{\rm hor}\rangle\bigg|_{s=0}
  =
  -2\pi
  \int\dd^{D-2}y\,\varphi(y)
  \int_{-\infty}^{\infty}\dd x^+\,
  \Delta_\Psi\langle T_{++}(x^+,0,y)\rangle,
  \label{eq:null-cut-full-modular-variation}
\end{equation}
where
\[
  \Delta_\Psi\langle X\rangle
  =
  \bra{\Psi}X\ket{\Psi}-\bra{\vac}X\ket{\vac}.
\]
If the chosen proof route supplies the half-sided modular inclusion
\(\mathfrak M_{s\varphi}\subset\mathfrak M_0\) for \(s>0\), with
\(\widehat K_0-\widehat K_{s\varphi}\) a positive quadratic form on the same
state class, then \eqref{eq:null-cut-full-modular-variation} gives
\begin{equation}
  \lim_{s\downarrow0}
  \frac{1}{s}
  \bra{\Psi}(\widehat K_0-\widehat K_{s\varphi})\ket{\Psi}
  =
  2\pi
  \int\dd^{D-2}y\,\varphi(y)
  \int_{-\infty}^{\infty}\dd x^+\,
  \Delta_\Psi\langle T_{++}(x^+,0,y)\rangle
  \geq0.
  \label{eq:anec-full-modular-inclusion-sign}
\end{equation}
The vacuum term vanishes by Poincare invariance and the stress-tensor
normalization.  This finite sign calculation explains why the null displacement
has the ANEC sign.  It is an independent proof only in a theorem package where
the null-cut modular formula, half-sided inclusion, common domain, and
differentiability statements have already been established without assuming
the desired inequality.

The perturbative deformed-half-space derivation is often written in finite
factor language as a relative-entropy squeeze between a cut and its
complement.  In a sharp continuum local algebra this is only a regulated
mnemonic.  To make that proof route valid one must specify a split or other
UV regulator, prove regulator independence of the Araki relative entropies and
modular shape derivatives, control the collar-removal limit, keep the state in
a common stress-tensor domain, and justify the equality of the region and
complement entropy first variations before the regulator is removed.  Once
those hypotheses are supplied, the right-cut derivative, the complementary
derivative, and their common regulated entropy variation reproduce the same
sign algebra as \eqref{eq:anec-full-modular-inclusion-sign}.

\paragraph{ANEC-to-detector bridge.}
The detector consequence contains a real limiting operation.  The route-scoped
ANEC input supplies positivity for smooth compact transverse tests on a null
plane.  The angular detector \(\mathcal E(f)\), by contrast, is a
null-infinity object tested by a smooth function
\(f\in C^\infty(S^{D-2})\), and finite calorimeter bins are obtained by
approximating indicator functions on the compact detector sphere.  To pass
from ANEC to the statewise detector measure
\(\mu_\Psi\) of \eqref{eq:cft-statewise-detector-measure}, one needs the
following additional bridge data:
\[
  f_j\geq0,\qquad f_j\to f\ \text{in }C(S^{D-2}),
  \qquad
  \bra{\Psi}\mathcal E(f_j)\ket{\Psi}
  \longrightarrow
  \bra{\Psi}\mathcal E(f)\ket{\Psi},
\]
with a bound controlled by the total energy of the state, so that the positive
quadratic form on smooth angular tests extends to a bounded positive
functional on \(C(S^{D-2})\).  Riesz--Markov then gives the finite positive
measure \(\mu_\Psi\).  Without this bounded-extension step, ANEC positivity
for each smooth transverse test is not yet a statement about Borel detector
bins, nor about products of detectors at separated or coincident angles.
Detector products require their own domain and contact-extension data, as in
the partition discussion following
\eqref{eq:finite-detector-partition-decomposition}.  This is why the chapter
keeps the one-detector positivity theorem, the statewise measure
construction, and the \(k\)-detector contact algebra as distinct pieces of
the detector framework.

The causality route packages the same sign in Lorentzian analyticity.  Let
\(\mathcal O\) be a scalar primary used only as a probe, and let
\(\mathcal X\) be a local operator whose wavepacket creates the finite-energy
state \(\Psi_{\mathcal X}\).  In the Regge light-cone regime of the normalized
four-point function
\[
  G=
  \frac{\bra{\vac}\mathcal O(x_1)\,\mathcal X^\dagger(x_3)
    \mathcal X(x_4)\,\mathcal O(x_2)\ket{\vac}}
  {\bra{\vac}\mathcal O(x_1)\mathcal O(x_2)\ket{\vac}
   \bra{\vac}\mathcal X^\dagger(x_3)\mathcal X(x_4)\ket{\vac}},
\]
the stress-tensor contribution to the first nontrivial discontinuity is
proportional to the ANEC matrix element
\[
  \bra{\Psi_{\mathcal X}}\mathcal A_n(\varphi)\ket{\Psi_{\mathcal X}},
  \qquad
  \varphi\geq0,
\]
where \(\varphi\) is determined by the probe wavepacket.  Reflection
positivity gives a positive norm on the Euclidean ordering, while Lorentzian
causality gives analyticity and a bounded phase in the continuation around
the light cone.  The maximum-modulus step in that analytic domain fixes the
sign of the stress-tensor discontinuity, hence gives
\eqref{eq:anec-positivity}.  The derivation is conditional on the same
interacting-theory, OPE-projection, convergence, analyticity,
order-of-limits, and infrared assumptions stated in
Quoted Theorem~\ref{qthm:anec-cft}; it is not the same theorem boundary as the
modular relative-entropy route.

For conformal collider physics, the consequence is immediate.  If a detector
state \(\Psi\) has total energy \(Q=\bra{\Psi}P^0\ket{\Psi}/\braket{\Psi}{\Psi}\)
and \(\mathcal E(\mathbf n)\) is the null-infinity representative of the
light transform, then
\[
  \frac{\bra{\Psi}\mathcal E(f)\ket{\Psi}}{\braket{\Psi}{\Psi}}\geq0
  \qquad (f\geq0),
  \qquad
  \int_{S^{D-2}}\dd\Omega(\mathbf n)\,
  \frac{\bra{\Psi}\mathcal E(\mathbf n)\ket{\Psi}}{\braket{\Psi}{\Psi}}
  =
  Q .
\]
Thus the angular energy profile in any fixed family of collider states is a
positive distribution on the celestial sphere with fixed total mass \(Q\).
The Hofman--Maldacena inequalities are the finite-dimensional specialization
of this positivity to stress-tensor states in four-dimensional CFTs.

\section{Conformal Collider One-Point Functions}

Consider a four-dimensional parity-even CFT and a stress-tensor state in the
rest frame \(q^\mu=(Q,\mathbf0)\).  Let \(\varepsilon_{ij}\) be a complex
symmetric traceless spatial polarization, \(i,j=1,2,3\), and write
\[
  \ket{\Psi_{\varepsilon,q}}
  =
  \varepsilon_{ij}\widetilde T^{ij}(q)\ket{\vac}
\]
with wavepacket regularization understood.  Rotational covariance around
\(\mathbf q=\mathbf0\), stress-tensor conservation, and the Ward identity
normalization of \(T_{\mu\nu}\) constrain the separated-angle one-detector
expectation to a two-parameter tensor structure.

\begin{lemma}[Rotational tensor basis for a stress-tensor collider state]
\label{lem:stress-tensor-collider-rotational-basis}
Let \(E_\varepsilon(\mathbf n)\) be a real quadratic form in a complex
symmetric traceless tensor \(\varepsilon_{ij}\), depending on
\(\mathbf n\in S^2\), and covariant under simultaneous spatial rotations of
\(\varepsilon\) and \(\mathbf n\).  Assume that its spherical-harmonic content
is obtained from the tensor product of two spin-two polarizations, as happens
for the separated-angle detector expectation in a stress-tensor state.  Then
after fixing the total-energy normalization, \(E_\varepsilon(\mathbf n)\) is
a linear combination of
\[
  1,\qquad
  I_2(\varepsilon,\mathbf n)
  =
  \frac{\varepsilon^*_{ij}\varepsilon_{ik}n^jn^k}
       {\varepsilon^*_{ij}\varepsilon_{ij}},
  \qquad
  I_4(\varepsilon,\mathbf n)
  =
  \frac{|\varepsilon_{ij}n^in^j|^2}
       {\varepsilon^*_{ij}\varepsilon_{ij}} .
\]
Their angular averages are
\begin{equation}
  \frac{1}{4\pi}\int_{S^2}\dd\Omega\, I_2=\frac13,
  \qquad
  \frac{1}{4\pi}\int_{S^2}\dd\Omega\, I_4=\frac{2}{15}.
  \label{eq:stress-collider-angular-averages}
\end{equation}
\end{lemma}

\begin{proof}
The bilinear expression in the polarization transforms in
\((2)\otimes(2)\) under \(SO(3)\), hence decomposes into angular momenta
\(\ell=0,1,2,3,4\).  Parity-even detector one-point functions and the
symmetry of the stress tensor keep only the even scalar structures.  The
\(\ell=0\) structure is the norm
\(\varepsilon^*_{ij}\varepsilon_{ij}\).  Contracting one pair of
polarization indices with \(n^jn^k\) gives the independent \(\ell=2\)
structure \(I_2\).  Contracting both pairs with \(n^in^j n^k n^\ell\) gives
the \(\ell=4\) structure \(I_4\) together with lower harmonics.  Thus
\(1,I_2,I_4\) span the parity-even rotational tensor basis.

The averages follow from the unique rotationally invariant tensors
\[
  \frac{1}{4\pi}\int_{S^2}\dd\Omega\, n^i n^j=\frac{\delta^{ij}}{3},
\]
and
\begin{equation}
  \frac{1}{4\pi}\int_{S^2}\dd\Omega\, n^i n^j n^k n^\ell
  =
  \frac{1}{15}
  \left(
    \delta^{ij}\delta^{k\ell}
    +\delta^{ik}\delta^{j\ell}
    +\delta^{i\ell}\delta^{jk}
  \right).
  \label{eq:s2-fourth-moment}
\end{equation}
The first identity gives
\[
  \langle I_2\rangle
  =
  \frac{\varepsilon^*_{ij}\varepsilon_{ik}\delta^{jk}}
       {3\,\varepsilon^*_{ab}\varepsilon_{ab}}
  =
  \frac13 .
\]
For \(I_4\), the term
\(\delta^{ij}\delta^{k\ell}\) in \eqref{eq:s2-fourth-moment} vanishes because
\(\varepsilon_{ij}\delta^{ij}=0\), while the remaining two contractions each
give \(\varepsilon^*_{ij}\varepsilon_{ij}\).  Hence
\(\langle I_4\rangle=2/15\).
\end{proof}

Consequently the separated-angle one-detector expectation has the form
\begin{equation}
  \frac{\bra{\Psi_{\varepsilon,q}}\mathcal E(\mathbf n)
  \ket{\Psi_{\varepsilon,q}}}
  {\braket{\Psi_{\varepsilon,q}}{\Psi_{\varepsilon,q}}}
  =
  \frac{Q}{4\pi}
  \left[
    1
    +t_2
    \left(
      \frac{\varepsilon^*_{ij}\varepsilon_{ik}n^jn^k}
           {\varepsilon^*_{ij}\varepsilon_{ij}}
      -\frac13
    \right)
    +t_4
    \left(
      \frac{|\varepsilon_{ij}n^in^j|^2}
           {\varepsilon^*_{ij}\varepsilon_{ij}}
      -\frac{2}{15}
    \right)
  \right].
  \label{eq:hm-energy-one-point}
\end{equation}
The real constants \(t_2,t_4\) are coordinates on the parity-even
stress-tensor three-point function after \(C_T\) and Ward-identity contact
terms have been fixed.  The subtractions \(1/3\) and \(2/15\) are precisely
the averages in \eqref{eq:stress-collider-angular-averages}, so integration
over \(S^2\) gives the total energy \(Q\) for every polarization.

\begin{lemma}[Spectral diagonalization of the stress-tensor collider form]
\label{lem:hm-helicity-spectral-diagonalization}
Fix the detector direction \(\mathbf n=e_3=(0,0,1)\), and let
\(\mathcal V\) be the complex vector space of symmetric traceless tensors
\(\varepsilon_{ij}\), equipped with
\(\|\varepsilon\|^2=\varepsilon^*_{ij}\varepsilon_{ij}\).  Decompose
\(\mathcal V\) under rotations around \(e_3\) as
\[
  \mathcal V=\mathcal V_2\oplus\mathcal V_1\oplus\mathcal V_0,
\]
where
\[
\begin{aligned}
  \mathcal V_2
  &=
  \{\varepsilon:\varepsilon_{13}=\varepsilon_{23}=\varepsilon_{33}=0\},\\
  \mathcal V_1
  &=
  \{\varepsilon:\varepsilon_{11}=\varepsilon_{22}
  =\varepsilon_{12}=\varepsilon_{33}=0\},\\
  \mathcal V_0
  &=
  \mathbb C\,\operatorname{diag}(1,1,-2).
\end{aligned}
\]
For the normal form \eqref{eq:hm-energy-one-point}, define the dimensionless
quadratic form
\[
  B_{t_2,t_4}(\varepsilon)
  =
  \|\varepsilon\|^2
  \left[
    1
    +t_2\left(I_2(\varepsilon,e_3)-\frac13\right)
    +t_4\left(I_4(\varepsilon,e_3)-\frac{2}{15}\right)
  \right].
\]
If
\(\varepsilon=\varepsilon_2+\varepsilon_1+\varepsilon_0\) is the above
orthogonal decomposition, then
\begin{equation}
  B_{t_2,t_4}(\varepsilon)
  =
  \lambda_2\|\varepsilon_2\|^2
  +\lambda_1\|\varepsilon_1\|^2
  +\lambda_0\|\varepsilon_0\|^2,
  \label{eq:hm-helicity-spectral-decomposition}
\end{equation}
with
\begin{equation}
  \lambda_2=1-\frac13t_2-\frac{2}{15}t_4,\qquad
  \lambda_1=1+\frac16t_2-\frac{2}{15}t_4,\qquad
  \lambda_0=1+\frac13t_2+\frac{8}{15}t_4 .
  \label{eq:hm-helicity-eigenvalues}
\end{equation}
\end{lemma}

\begin{proof}
The three displayed subspaces are pairwise orthogonal for
\(\varepsilon^*_{ij}\eta_{ij}\).  They exhaust \(\mathcal V\): a symmetric
traceless \(3\times3\) tensor is determined by the two transverse traceless
entries \((\varepsilon_{11}-\varepsilon_{22},\varepsilon_{12})\), the two
mixed entries \((\varepsilon_{13},\varepsilon_{23})\), and the scalar trace-free
diagonal direction proportional to \(\operatorname{diag}(1,1,-2)\).

For \(\varepsilon_2\in\mathcal V_2\), the vector
\(\varepsilon_2 e_3\) and the scalar \(e_3^i\varepsilon_{2,ij}e_3^j\) both
vanish, so \(I_2=I_4=0\).  For
\(\varepsilon_1\in\mathcal V_1\), writing
\(\varepsilon_{13}=a\) and \(\varepsilon_{23}=b\), one has
\[
  \|\varepsilon_1\|^2=2(|a|^2+|b|^2),\qquad
  \|\varepsilon_1 e_3\|^2=|a|^2+|b|^2,\qquad
  e_3^i\varepsilon_{1,ij}e_3^j=0,
\]
so \(I_2=1/2\) and \(I_4=0\).  For
\(\varepsilon_0=c\,\operatorname{diag}(1,1,-2)\), one has
\[
  \|\varepsilon_0\|^2=6|c|^2,\qquad
  \|\varepsilon_0 e_3\|^2=4|c|^2,\qquad
  |e_3^i\varepsilon_{0,ij}e_3^j|^2=4|c|^2,
\]
so \(I_2=I_4=2/3\).  These three evaluations give the eigenvalues
\eqref{eq:hm-helicity-eigenvalues}.  The mixed terms vanish because the
subspaces are orthogonal and because the numerators
\(\|\varepsilon e_3\|^2\) and
\(|e_3^i\varepsilon_{ij}e_3^j|^2\) separate into the same
\(\mathcal V_2,\mathcal V_1,\mathcal V_0\) components.  This proves
\eqref{eq:hm-helicity-spectral-decomposition}.
\end{proof}

\paragraph{Four-dimensional conformal-collider inequalities.}
The finite collider step begins after two analytic ingredients have already
been supplied: the ANEC-to-detector bridge that makes
\(\mathcal E(f)\) a positive angular functional for \(f\geq0\), and the
Wightman/Ward calculation that puts the stress-tensor one-point function in
the normal form \eqref{eq:hm-energy-one-point}.  The theorem-level content
inside the remaining finite step is
Lemma~\ref{lem:hm-helicity-spectral-diagonalization}: it identifies the three
irreducible \(SO(2)\) polarization sectors and computes the corresponding
eigenvalues of the detector quadratic form.  The fixed-total-energy identity
controls only the zeroth spherical moment.  Positivity is the sharper
statement that every allowed polarization gives a nonnegative detector density
in every direction, so the normal form \eqref{eq:hm-energy-one-point} is
positive for every stress-tensor polarization and every detector direction
\(\mathbf n\) exactly when
\[
  1-\frac13t_2-\frac{2}{15}t_4\geq0,
  \qquad
  1+\frac16t_2-\frac{2}{15}t_4\geq0,
  \qquad
  1+\frac13t_2+\frac{8}{15}t_4\geq0.
\]
Indeed, by rotational covariance one may put \(\mathbf n=e_3\); the lemma
then writes the form as
\(\lambda_2\|\varepsilon_2\|^2+\lambda_1\|\varepsilon_1\|^2+
\lambda_0\|\varepsilon_0\|^2\), and rotation transports the same three
eigenvalues to any direction.  Each of the three helicity sectors is
independent in this test: fixing the average energy and checking only two
sectors still leaves normalized stress-tensor profiles whose omitted sector
has negative flux.

The derivation of the normal form
\eqref{eq:hm-energy-one-point} is a calculation of a Wightman
\(\langle TTT\rangle\) distribution integrated against the detector kernel.
Contact terms in the local \(\langle TTT\rangle\) Ward identities fix the
normalization of the stress tensor and do not change separated-angle detector
values after the state wavepacket and detector smearing have been specified.

\paragraph{Supersymmetric central-charge specialization.}
In a four-dimensional \(\mathcal N=1\) SCFT the stress tensor belongs to the
supercurrent multiplet, so supersymmetric Ward identities reduce the
parity-even stress-tensor three-point data to the two trace-anomaly
coefficients \(a\) and \(c\).  In the collider normalization used in
\eqref{eq:hm-energy-one-point}, the separated-angle stress-tensor flux
coordinates are
\begin{equation}
  t_4=0,\qquad
  t_2=6\left(1-\frac{a}{c}\right).
  \label{eq:n1-susy-collider-central-charge-map}
\end{equation}
Here \(c>0\) is fixed by the stress-tensor two-point normalization.  Substituting
\eqref{eq:n1-susy-collider-central-charge-map} into the three helicity
eigenvalues \eqref{eq:hm-helicity-eigenvalues} gives
\begin{equation}
  \lambda_2=2\frac{a}{c}-1,\qquad
  \lambda_1=2-\frac{a}{c},\qquad
  \lambda_0=3-2\frac{a}{c}.
  \label{eq:n1-susy-collider-helicity-eigenvalues}
\end{equation}
The middle inequality is then weaker than the two endpoint inequalities, and
energy-detector positivity gives the familiar supersymmetric collider bound
\begin{equation}
  \frac12\leq\frac{a}{c}\leq\frac32 .
  \label{eq:n1-susy-collider-ac-bound}
\end{equation}
This is the point at which the abstract detector positivity becomes a
constraint on physical anomaly data.  The free chiral multiplet has
\((a,c)=(1/48,1/24)\), hence \(a/c=1/2\) and saturates the helicity-two
inequality.  The free vector multiplet has
\((a,c)=(3/16,1/8)\), hence \(a/c=3/2\) and saturates the helicity-zero
inequality.  Interacting theories need not sit at either endpoint, but they
must transport the same stress-tensor normalization and supercurrent Ward
identity before \eqref{eq:n1-susy-collider-ac-bound} can be compared across
dual descriptions.

\section{Energy-Energy Correlators in a CFT}

For a state of sharp total energy \(Q\), the normalized energy-energy
correlator is the distribution
\begin{equation}
  \mathrm{EEC}_\Psi(\chi)
  =
  \frac{1}{Q^2\braket{\Psi}{\Psi}}
  \bra{\Psi}
  \int_{S^{D-2}}\dd\Omega_1\dd\Omega_2\,
  \delta(\cos\chi-\mathbf n_1\cdot\mathbf n_2)
  \mathcal E(\mathbf n_1)\mathcal E(\mathbf n_2)
  \ket{\Psi}.
  \label{eq:cft-eec}
\end{equation}
For a wavepacket in \(Q\), replace \(Q^2\) by the corresponding expectation
of \((P^0)^2\), or keep the unnormalized numerator and denominator separately.
The separated-angle CFT observable
\eqref{eq:cft-eec} is the same type of detector distribution as the QCD
observable \eqref{eq:qcd-eec-definition}.  The difference is dynamical:
in a CFT there is no mass gap and no hadronic scattering basis, while the
detector remains defined by the stress tensor and null infinity.

When two detector directions approach one another, the product of light-ray
operators has a short-distance expansion on the celestial sphere.  The
following theorem is used as an external theorem boundary for the modern
light-ray formalism.\footnote{See M.~Kologlu, P.~Kravchuk,
D.~Simmons-Duffin, and A.~Zhiboedov, ``The light-ray OPE and conformal
colliders,'' arXiv:1905.01311.  For the conformal-collider formulation, see
D.~M.~Hofman and J.~Maldacena, ``Conformal collider physics: energy and charge
correlations,'' JHEP 05 (2008) 012, arXiv:0803.1467.}

\begin{quotedtheorem}[Light-ray OPE under Lorentzian CFT hypotheses]
\label{qthm:light-ray-ope-boundary}
In a unitary CFT satisfying the Lorentzian analyticity, spectrum, and
operator-growth hypotheses required for the light-ray OPE, products of
null-integrated operators on the same null plane admit a convergent expansion
in light-ray operators.  For energy detectors, the separated-angle product has
the form, as a distribution tested near \(\mathbf n_1=\mathbf n_2\),
\[
  \mathcal E(\mathbf n_1)\mathcal E(\mathbf n_2)
  =
  \sum_\alpha
  C_\alpha(\mathbf n_1,\mathbf n_2)\,
  \mathbb L_\alpha(\mathbf n_2),
\]
where \(\mathbb L_\alpha\) are light-ray transforms of local CFT primaries and
their descendants, analytically continued in spin as specified by the
Lorentzian inversion formula.
\end{quotedtheorem}

The word ``convergent'' in
Quoted Theorem~\ref{qthm:light-ray-ope-boundary} is a distributional
statement, not a pointwise multiplication rule for two unbounded operators.
On a small celestial patch write a detector direction as a transverse
coordinate \(y\in\mathbb R^{D-2}\).  A separated detector product is first
paired with a test function
\[
  \Phi(y_1,y_2)
  \in C_c^\infty\bigl((U\times U)\setminus \{y_1=y_2\}\bigr).
\]
The diagonal limit is probed by families whose support approaches
\(|y_{12}|=|y_1-y_2|=0\).  The theorem supplies coefficient distributions
\[
  C_{\alpha,m}(y_{12})
\]
and light-ray operators \(\mathbb L_\alpha(y)\), together with a topology of
test functions and state wavepackets in which
\begin{equation}
  \mathcal E(y_1)\mathcal E(y_2)
  \sim
  \sum_{\alpha,m}
  C_{\alpha,m}(y_{12})\,
  \partial_y^m\mathbb L_\alpha(y_2)
  \label{eq:cft-light-ray-ope-local-form}
\end{equation}
after pairing with such tests.  The multi-index \(m\) labels transverse
descendants and tensor structures.  The coefficients are homogeneous or
polyhomogeneous distributions away from \(y_{12}=0\), with logarithmic terms
when anomalous dimensions or degeneracies are present.  Their normalization
is fixed by three-point data involving the two detector light transforms and
\(\mathbb L_\alpha\), or equivalently by the Lorentzian inversion transform
in the conformal frame used to define the light transform.

The proof mechanism has three distinct analytic steps.  First, the product of
the local stress tensors is not expanded by the Euclidean local OPE at
ordinary spacelike separation.  The two null integrations and the detector
limit force one to control Lorentzian configurations in which the integration
contours approach lightlike singularities.  Lorentzian analyticity and the
spectrum condition provide domains in complexified configuration space in
which the contours can be displaced and the relevant discontinuities can be
defined.  Second, Regge or operator-growth bounds control the arcs at
infinity and justify exchanging the null-line integrals with the partial-wave
or inversion expansion.  Without these bounds the formal operation
``integrate the local OPE along a null line'' is not a proof.  Third, the
resulting expansion is completed in a transverse distribution topology.  A
usable version states that, after retaining all light-ray primaries in a set
\(\mathcal A_N\), the remainder \(R_N\) obeys an estimate of the form
\begin{equation}
  \bigl|
  \bra{\Psi}R_N(\Phi)\ket{\Psi}
  \bigr|
  \leq
  C_N(\Psi)\,
  \|\Phi\|_{\mathcal T_N}\,
  \epsilon^{\sigma_N}
  \label{eq:cft-light-ray-ope-remainder-template}
\end{equation}
for tests supported in \(|y_{12}|<\epsilon\), with a declared test-function
seminorm \(\|\cdot\|_{\mathcal T_N}\) and an exponent \(\sigma_N>0\) that
improves as the retained operator set is enlarged.  Equation
\eqref{eq:cft-light-ray-ope-remainder-template} is not an additional theorem
proved here; it is the kind of bound that makes
\eqref{eq:cft-light-ray-ope-local-form} a mathematical statement rather than
a formal series.  Whenever the monograph uses the light-ray OPE below, the
use is restricted to consequences that follow from such Lorentzian
analyticity, growth, and transverse-distribution convergence hypotheses.

\paragraph{Finite light-ray OPE charts.}
For calculations and for comparison with detector data, one never manipulates
the entire light-ray OPE at once.  One chooses a finite chart and states its
remainder.  Work in a small angular patch with transverse coordinates
\((y_1,y_2)\), relative coordinate \(\xi=y_1-y_2\), and center coordinate
\(y=y_2\).  Let \(\mathfrak A_N\) be a finite set of retained light-ray
primary and descendant labels \((\alpha,m)\).  A finite CFT light-ray OPE
chart for the two-detector product in a state domain \(\mathcal D_N\)
consists of:
\begin{enumerate}
  \item coefficient distributions \(C_{\alpha,m}\) in the relative variable
  \(\xi\), including their specified extension across \(\xi=0\) whenever
  contact terms are retained;
  \item light-ray quadratic forms
  \(\mathbb L_\alpha(\psi)\) on \(\mathcal D_N\), where \(\psi\) is a smooth
  compactly supported center-coordinate test;
  \item for each test
  \(\Phi\in C_c^\infty(U\times U)\), the induced center tests
  \begin{equation}
    \psi_{\alpha,m,\Phi}(y)
    =
    \bigl\langle C_{\alpha,m}(\xi),
    \partial_y^m\Phi(y+\xi,y)\bigr\rangle_{\xi};
    \label{eq:cft-finite-light-ray-chart-center-test}
  \end{equation}
  \item a seminorm \(\|\cdot\|_{\mathcal T_N}\), a neighborhood condition on
  the support of \(\Phi\), and a remainder functional
  \(R_N(\Phi;\Psi)\) such that
  \begin{equation}
    \left|
      R_N(\Phi;\Psi)
    \right|
    \leq
    C_N(\Psi)\,\|\Phi\|_{\mathcal T_N}\,\epsilon^{\sigma_N},
    \qquad \sigma_N>0,
    \label{eq:cft-finite-light-ray-chart-remainder}
  \end{equation}
  for tests supported in \(|\xi|<\epsilon\).
\end{enumerate}
The finite chart represents the detector product by
\begin{equation}
  \bra{\Psi}\mathcal E\mathcal E(\Phi)\ket{\Psi}
  =
  \sum_{(\alpha,m)\in\mathfrak A_N}
  \bra{\Psi}\mathbb L_\alpha(\psi_{\alpha,m,\Phi})\ket{\Psi}
  +
  R_N(\Phi;\Psi).
  \label{eq:cft-finite-light-ray-chart-expansion}
\end{equation}
Here \(\mathcal E\mathcal E(\Phi)\) denotes the distributional pairing of the
two-detector product with \(\Phi\), including the declared contact convention.

This finite chart is the CFT analogue of the QCD finite renormalized
light-ray mixing chart in
Chapter~\ref{ch:volume-ii-19}, but the
coordinates are fixed-point conformal data rather than scale-dependent
renormalized QCD operators.  The coefficient extension across \(\xi=0\) is
part of the chart.  A separated-angle chart that only tests
\(\Phi\) supported away from the diagonal does not determine the coincident
detector contact coordinate, hence does not determine the full moment
\(\mathcal G_2(1,1;\Psi)=\langle\Psi|(P^0)^2|\Psi\rangle\).

For a fixed state \(\Psi\), suppose the retained light-ray forms and
coefficient maps obey
\begin{equation}
  \left|
    \bra{\Psi}\mathbb L_\alpha(\psi)\ket{\Psi}
  \right|
  \leq
  L_{\alpha}(\Psi)\,p_\alpha(\psi),
  \qquad
  p_\alpha(\psi_{\alpha,m,\Phi})
  \leq
  K_{\alpha,m}\,\|\Phi\|_{\mathcal T_N}.
  \label{eq:cft-finite-light-ray-chart-continuity}
\end{equation}
Then \eqref{eq:cft-finite-light-ray-chart-expansion} gives the finite
detector-functional bound
\begin{equation}
  \left|
    \bra{\Psi}\mathcal E\mathcal E(\Phi)\ket{\Psi}
  \right|
  \leq
  \left(
    \sum_{(\alpha,m)\in\mathfrak A_N}
    L_\alpha(\Psi)K_{\alpha,m}
    +
    C_N(\Psi)\epsilon^{\sigma_N}
  \right)
  \|\Phi\|_{\mathcal T_N}.
  \label{eq:cft-finite-light-ray-chart-detector-bound}
\end{equation}
Thus the finite light-ray chart consists of coefficient distributions, their
diagonal extensions, continuity estimates for the light-ray forms, and a
remainder bound in the detector-test topology.  These data are what turn a
truncated representation-theoretic expansion into a controlled approximation
to the energy correlator.

\paragraph{Finite chart changes and contact coordinates.}
The retained light-ray basis and the diagonal extension convention are part of
the finite chart.  Write the retained contribution to
\eqref{eq:cft-finite-light-ray-chart-expansion} in finite-dimensional notation
as
\[
  V_\Psi(\Phi)
  =
  c_\Phi\,\ell_\Psi+k_\Phi+R_N(\Phi;\Psi).
\]
Here \(\ell_\Psi\in\mathbb C^r\) is the column vector of retained light-ray
matrix elements, \(c_\Phi\in(\mathbb C^r)^\ast\) is the row vector obtained by
pairing the coefficient distributions with \(\Phi\), and \(k_\Phi\) denotes
the explicit diagonal-contact coordinate in the chosen extension.  If
\(B\in GL(r,\mathbb C)\) changes the retained light-ray basis, the same
detector functional is represented by
\begin{equation}
  \ell'_\Psi=B^{-1}\ell_\Psi,
  \qquad
  c'_\Phi=c_\Phi B,
  \qquad
  c'_\Phi\ell'_\Psi=c_\Phi\ell_\Psi .
  \label{eq:cft-light-ray-chart-basis-covariance}
\end{equation}
This is ordinary coordinate covariance of the retained finite vector space;
it is not a dynamical statement about the exact OPE.

Changing the extension of the relative-coordinate coefficient distributions
across the diagonal is different.  Let \(d_\Phi\in(\mathbb C^r)^\ast\) be the
row vector produced by adding diagonal distributions to the retained
coefficient maps.  The full detector distribution is unchanged only if the
explicit contact coordinate is shifted simultaneously:
\begin{equation}
  c_\Phi^{\rm new}=c_\Phi+d_\Phi,
  \qquad
  k_\Phi^{\rm new}=k_\Phi-d_\Phi\ell_\Psi .
  \label{eq:cft-light-ray-chart-contact-reshuffle}
\end{equation}
Thus a separated-angle calculation, for which \(d_\Phi=0\) on all tests used,
does not determine \(k_\Phi\).  The Ward identity
\(\mathcal G_2(1,1;\Psi)=\langle\Psi|(P^0)^2|\Psi\rangle\) fixes a condition
on the sum \(c_\Phi\ell_\Psi+k_\Phi\) for the constant detector test; it does
not fix a separated-angle coefficient convention by itself.  A finite
light-ray OPE chart used for a full energy correlator must therefore state
both its retained-basis convention and its diagonal-contact convention.

\paragraph{One-parameter finite chart transport.}
The same bookkeeping can be differentiated.  This is useful because it
separates a harmless change of finite coordinates from a genuine change of the
detector distribution.  Let \(s\) parametrize a family of retained-basis
changes \(B_s\in GL(r,\mathbb C)\) and diagonal-extension shifts
\(d_{s,\Phi}\in(\mathbb C^r)^*\), both stated in the original retained
coordinates.  For
\[
  V_\Psi(\Phi)=c_\Phi\ell_\Psi+k_\Phi+R_N(\Phi;\Psi)
\]
define
\begin{equation}
  \ell_{s,\Psi}=B_s^{-1}\ell_\Psi,\qquad
  c_{s,\Phi}=(c_\Phi+d_{s,\Phi})B_s,\qquad
  k_{s,\Phi}=k_\Phi-d_{s,\Phi}\ell_\Psi .
  \label{eq:cft-light-ray-chart-transport-coordinates}
\end{equation}
Then
\[
  c_{s,\Phi}\ell_{s,\Psi}+k_{s,\Phi}
  =
  c_\Phi\ell_\Psi+k_\Phi .
\]
Put \(A_s=B_s^{-1}\partial_sB_s\).  Differentiating
\eqref{eq:cft-light-ray-chart-transport-coordinates} gives
\begin{align}
  \partial_s\ell_{s,\Psi}
  &=
  -A_s\ell_{s,\Psi},&
  \partial_sc_{s,\Phi}
  &=
  c_{s,\Phi}A_s+(\partial_sd_{s,\Phi})B_s,
  \label{eq:cft-light-ray-chart-transport-row-column}\\
  \partial_sk_{s,\Phi}
  &=
  -(\partial_sd_{s,\Phi})B_s\ell_{s,\Psi}.
  \label{eq:cft-light-ray-chart-transport-contact}
\end{align}
The \(c_{s,\Phi}A_s\) and \(-A_s\ell_{s,\Psi}\) terms cancel in
\(\partial_s(c_s\ell_s+k_s)\); the remaining
\((\partial_sd_s)B_s\ell_s\) term cancels only against the contact derivative
\eqref{eq:cft-light-ray-chart-transport-contact}.  With the opposite contact
sign the constant detector test would change by twice this endpoint
contribution.  Equations
\eqref{eq:cft-light-ray-chart-transport-row-column}--\eqref{eq:cft-light-ray-chart-transport-contact}
are finite-chart identities at a conformal fixed point, not anomalous
dimensions.  Their QCD analogue is the renormalized light-ray mixing chart of
Chapter~\ref{ch:volume-ii-19}, where \(s\) is replaced by physical
renormalization or rapidity scales after the regulator, subtraction scheme,
and endpoint contact convention have been fixed.

\paragraph{Small-angle pushforward to the EEC variable.}
The finite light-ray chart lives first as a distribution in the two detector
directions.  The one-variable EEC is obtained by pushing that distribution
forward by the opening angle.  This geometric step contributes a Jacobian that
must be included together with the light-ray OPE coefficient.  Put
\[
  d_\perp=D-2
\]
for the dimension of the detector sphere, fix the second detector direction,
and write \(\theta\) for the geodesic distance of the first detector from it.
In geodesic polar coordinates on \(S^{d_\perp}\),
\[
  \dd\Omega_1=(\sin\theta)^{d_\perp-1}\dd\theta\,\dd\omega,
  \qquad \omega\in S^{d_\perp-1}.
\]
The EEC kernel in \eqref{eq:cft-eec} is
\(\delta(\cos\chi-\cos\theta)\), hence for \(0<\chi<\pi\)
\begin{equation}
  \int_{S^{d_\perp}}\dd\Omega_1\,
  \delta(\cos\chi-\mathbf n_1\cdot\mathbf n_2)\,
  F(\theta,\omega)
  =
  (\sin\chi)^{d_\perp-2}
  \int_{S^{d_\perp-1}}\dd\omega\,F(\chi,\omega).
  \label{eq:cft-eec-small-angle-pushforward}
\end{equation}
Thus a retained coefficient whose relative-coordinate leading term is
homogeneous of degree \(-\lambda\),
\[
  C(\xi)=|\xi|^{-\lambda}C_0(\widehat\xi)+\text{higher radial orders},
  \qquad \xi=\theta\,\omega ,
\]
contributes to the small-angle EEC density with leading power
\begin{equation}
  \mathrm{EEC}_C(\chi)
  \sim
  \chi^{d_\perp-2-\lambda}
  \int_{S^{d_\perp-1}}\dd\omega\,C_0(\omega)\,
  \langle\Psi|\mathbb L(\psi_\omega)|\Psi\rangle ,
  \qquad \chi\downarrow0,
  \label{eq:cft-eec-small-angle-homogeneous-power}
\end{equation}
up to the center-coordinate smearing and the declared finite-chart remainder.
Polyhomogeneous terms
\(|\xi|^{-\lambda_a}(\log|\xi|)^s C_{a,s}(\widehat\xi)\) produce the
corresponding powers
\(\chi^{d_\perp-2-\lambda_a}(\log\chi)^s\), with curvature corrections from
\(\sin\chi/\chi\) entering at even higher powers.  If one changes from
\(\chi\) to \(z=(1-\cos\chi)/2\), the leading power is divided by two because
\(z=\chi^2/4+O(\chi^4)\).  Therefore the exponent quoted for a small-angle
EEC singularity is a statement about the light-ray coefficient homogeneity,
the detector-sphere dimension, and the chosen one-variable kernel.

This formulation also explains the relation with the QCD small-angle EEC.
In a CFT the expansion coefficients are fixed-point data of light-ray
representations.  In asymptotically free QCD, the corresponding small-angle
statement is a renormalized collinear light-ray expansion with a scale
\(\mu\), an anomalous-dimension matrix, endpoint contact coordinates, and a
mass-gap-sensitive long-distance remainder.  The QCD object is therefore not
obtained by simply copying the CFT theorem; it is a separate factorization
datum whose detector side is the same stress-tensor observable.

\paragraph{Endpoint distribution charts in the EEC variable.}
The pushforward to one angular variable is still not the final endpoint
distribution.  A small-angle light-ray coefficient gives a density on
\(z>0\), where
\[
  z=\frac{1-\cos\chi}{2},
  \qquad z=0
\]
is the coincident-detector endpoint.  To compare with the full detector
distribution, the density must be extended across \(z=0\) together with an
explicit contact coordinate.  This is the same logical distinction as in the
two-detector chart above: separated-angle data determine a distribution on
\((0,z_0)\), while the full detector product also contains diagonal contact
strata.

Fix an endpoint interval \(0<z<z_0\).  Let \(f(z)\) be one singular density
appearing in a finite light-ray chart, and suppose that subtracting the Taylor
jet of order \(J\) of the test function makes the integral finite.  The
subtracted endpoint extension is the distribution
\begin{equation}
  \left\langle [f]_{J,z_0,+},\varphi\right\rangle
  :=
  \int_0^{z_0} f(z)
  \left(
    \varphi(z)
    -
    \sum_{j=0}^J
      \frac{z^j}{j!}\varphi^{(j)}(0)
  \right)\dd z .
  \label{eq:cft-eec-subtracted-endpoint-extension}
\end{equation}
For the logarithmic \(1/z\) endpoint singularity this reduces to the familiar
plus prescription
\[
  \left\langle
    \left[\frac{\Theta(z_0-z)}{z}\right]_{+},
    \varphi
  \right\rangle
  =
  \int_0^{z_0}
    \frac{\varphi(z)-\varphi(0)}{z}\,\dd z .
\]
Terms more singular than this are not hidden in the word ``plus'': they
require the corresponding higher Taylor subtraction and explicit
\(\delta^{(j)}(z)\) contact coordinates, with
\(\langle\delta^{(j)},\varphi\rangle=(-1)^j\varphi^{(j)}(0)\).

A finite endpoint chart for the CFT EEC therefore has the form
\begin{equation}
  \mathrm{EEC}^{(N)}_\Psi
  =
  \mathcal R^{(N)}_{z_0,\Psi}
  +
  \sum_{a\in A_N}
    c_a(\Psi)\,[f_a]_{J_a,z_0,+}
  +
  \sum_{j=0}^{J_N}
    K_j^{(N)}(\Psi)\,\delta^{(j)}(z)
  +
  R^{(N)}_{z_0,\Psi}.
  \label{eq:cft-eec-finite-endpoint-chart}
\end{equation}
Here \(\mathcal R^{(N)}_{z_0,\Psi}\) is the ordinary distributional bulk
representative away from \(z=0\), the \(c_a(\Psi)\) are the retained
light-ray matrix elements and coefficient normalizations after the angular
pushforward, and \(R^{(N)}_{z_0,\Psi}\) is the declared finite-chart
remainder.  The coefficients \(K_j^{(N)}\) are not determined by knowing the
separated-angle asymptotic densities \(f_a\) alone.  They are fixed only after
the diagonal detector extension, Ward identities, and endpoint matching
convention have all been specified.  In particular, the constant-detector
moment fixes a linear combination of the separated-angle integral and the
contact coordinates; it does not by itself choose the representative in
\eqref{eq:cft-eec-finite-endpoint-chart}.

The dependence on \(z_0\) is a useful sign check.  For the basic plus
distribution
\[
  D_{z_*}:=
  \left[\frac{\Theta(z_*-z)}{z}\right]_{+}
\]
one has, as distributions on \([0,z_b]\),
\begin{equation}
  D_{z_b}
  =
  D_{z_a}
  +
  \frac{\mathbf 1_{z_a<z<z_b}}{z}
  -
  \log\frac{z_b}{z_a}\,\delta(z),
  \qquad 0<z_a<z_b .
  \label{eq:cft-eec-endpoint-resolution-shift}
\end{equation}
Testing \eqref{eq:cft-eec-endpoint-resolution-shift} against \(\varphi\)
gives exactly
\[
  \int_0^{z_a}\frac{\varphi-\varphi(0)}{z}\dd z
  +
  \int_{z_a}^{z_b}\frac{\varphi(z)}{z}\dd z
  -
  \log\frac{z_b}{z_a}\varphi(0)
  =
  \int_0^{z_b}\frac{\varphi-\varphi(0)}{z}\dd z .
\]
Equivalently, if the ordinary annulus \(z_a<z<z_b\) is moved from the bulk
representative into the endpoint plus-chart, the explicit \(\delta(z)\)
coordinate must shift by the annulus constant
\(\log(z_b/z_a)\) times the coefficient of \(1/z\).  With the opposite sign,
the constant test changes, so the total-energy Ward identity would fail.  The
QCD endpoint formula
\eqref{eq:qcd-eec-plus-resolution-shift} is the running-coupling,
scheme-dependent analogue of this fixed-point distributional identity.

This endpoint chart gives a nonperturbative meaning to the collinear
expansion of energy correlators in a CFT.  It is the conformal counterpart of
the QCD statement that small-angle energy correlators are governed by
collinear operator data and anomalous dimensions.  In QCD the running coupling
and the mass gap alter the long-distance behavior; in a CFT the light-ray OPE
is an operator expansion intrinsic to the fixed point, supplemented by the
distributional contact data needed to recover the full detector product.

\section{Comparison with QCD Detector Observables}

The QCD and CFT constructions share the same operator-theoretic skeleton:
\[
  T^{\mu\nu}
  \quad\leadsto\quad
  \mathcal E(\mathbf n)
  \quad\leadsto\quad
  \mathcal G_k(\mathbf n_1,\ldots,\mathbf n_k).
\]
In QCD, the physical final-state resolution is a hadronic spectral sum and
perturbation theory enters through asymptotically free short-distance
coefficient functions.  In a CFT, the state can be created directly by local
operators and the angular dependence is constrained by conformal
representation theory.  The common lesson is mathematical rather than
phenomenological: a detector observable is defined by the stress tensor in the
physical Hilbert space, and partonic or conformal expansions are calculational
representations of that observable under stated hypotheses.
