\chapter{Primary Operators and Finite Conformal Transformations}
\label{ch:volume-iii-primary-operators-finite-transformations}
\ConventionsEuclideanCFT


\section{Spin Representations and Primary Data}

\begin{definition}[Spin primary datum]
\label{def:spin-primary-datum}
Let \(\rho\) be a finite-dimensional representation of \(SO(D)\) on a vector
space \(V_\rho\).  A local operator valued in this representation is written
\[
  \mathcal O_a(x),
  \qquad
  a=1,\ldots,\dim V_\rho .
\]
The infinitesimal spin matrices are denoted
\[
  (S_{\mu\nu})_a{}^b=-(S_{\nu\mu})_a{}^b .
\]
They obey the \(SO(D)\) Lie algebra in the representation \(\rho\), in the
convention in which a finite rotation \(R\) near the identity is represented as
\(\rho(R)=\exp(-\frac{\ii}{2}\omega^{\mu\nu}S_{\mu\nu})\).  Thus the
\(S_{\mu\nu}\) are Hermitian spin matrices in a unitary representation.

A conformal primary operator of spin \(\rho\) and scaling dimension
\(\Delta\) is an operator whose state at the origin obeys
\begin{equation}
  \widehat K_\mu\ket{\mathcal O_a}=0,\qquad
  \widehat D_{\rm rad}\ket{\mathcal O_a}=\Delta\ket{\mathcal O_a},\qquad
  \widehat J_{\mu\nu}\ket{\mathcal O_a}
  =
  -(S_{\mu\nu})_a{}^b\ket{\mathcal O_b}.
  \label{eq:primary-state-data}
\end{equation}
\end{definition}

\begin{remark}[Global primaries, quasi-primaries, and Virasoro primaries]
\label{rem:global-primary-quasiprimary-terminology}
In \(D\ge3\) the conformal Killing equation on flat space has only the
finite-dimensional conformal algebra.  Therefore the state condition
\eqref{eq:primary-state-data} is the intrinsic notion of a conformal primary
used throughout this volume.  In \(D=2\) there is an additional local
conformal enhancement: on the plane the globally defined conformal group is
\(\mathrm{PSL}(2,\mathbb C)\), generated by
\(L_{-1},L_0,L_1\) and \(\bar L_{-1},\bar L_0,\bar L_1\), while the local
holomorphic and antiholomorphic conformal transformations are represented
quantum mechanically by the Virasoro algebras.  A local operator whose state
is annihilated by \(L_1\) and \(\bar L_1\) is often called a
\emph{quasi-primary}; it transforms covariantly under the global conformal
group.  A Virasoro primary is a stronger object: its state is annihilated by
all \(L_n,\bar L_n\) with \(n>0\), after choosing the stress-tensor and
Virasoro representation.  Thus, whenever this core volume uses the word
primary in \(D=2\) without further qualification, it means global primary in
the preceding sense, i.e. quasi-primary in much of the two-dimensional CFT
literature.  A later treatment that uses full Virasoro symmetry must state
the stronger Virasoro-primary condition separately.
\end{remark}

The translation generators create the descendant states:
\[
  \widehat P_{\mu_1}\cdots\widehat P_{\mu_k}\ket{\mathcal O_a}
  \longleftrightarrow
  \partial_{\mu_1}\cdots\partial_{\mu_k}\mathcal O_a(0).
\]

\begin{definition}[Conformal generalized Verma module]
\label{def:conformal-generalized-verma-module}
Let
\[
  \mathfrak g_{\mathbb C}
  =
  \mathfrak{so}(D+2,\mathbb C)
\]
be the complexified conformal Lie algebra, written in the radial triangular
decomposition
\[
  \mathfrak g_{\mathbb C}
  =
  \mathfrak n_+
  \oplus
  \mathfrak l
  \oplus
  \mathfrak n_-,
  \qquad
  \mathfrak n_+=\operatorname{span}_{\mathbb C}\{\widehat P_\mu\},
  \qquad
  \mathfrak n_-=\operatorname{span}_{\mathbb C}\{\widehat K_\mu\},
\]
\[
  \mathfrak l
  =
  \mathfrak{so}(D,\mathbb C)
  \oplus
  \mathbb C\,\widehat D_{\rm rad}.
\]
Let \(V_{\Delta,\rho}\) be the finite-dimensional
\(\mathfrak l\oplus\mathfrak n_-\)-module on which
\(\widehat D_{\rm rad}\) acts by \(\Delta\), \(\mathfrak{so}(D)\) acts by
\(\rho\), and every \(\widehat K_\mu\) acts by zero.  The conformal
generalized Verma module generated by this primary datum is the induced module
\begin{equation}
  M(\Delta,\rho)
  =
  U(\mathfrak g_{\mathbb C})
  \otimes_{U(\mathfrak l\oplus\mathfrak n_-)}
  V_{\Delta,\rho}.
  \label{eq:conformal-generalized-verma-module}
\end{equation}
By the Poincare--Birkhoff--Witt theorem and the commutativity of the
translations, its level-\(n\) vector space is
\begin{equation}
  M(\Delta,\rho)_n
  \simeq
  \operatorname{Sym}^n(\mathbb C^D)\otimes V_\rho,
  \qquad n\ge0,
  \label{eq:conformal-verma-levels}
\end{equation}
before equations of motion, conservation laws, and null relations are imposed.
\end{definition}

\begin{proposition}[Conformal multiplet generated by a primary datum]
\label{prop:conformal-multiplet-from-primary-datum}
Thus, under the state--operator correspondence, a positive-energy conformal
module is generated from a lowest-energy state by repeated action of
\(\widehat P_\mu\); algebraically this is the generalized Verma module
\(M(\Delta,\rho)\).  A descendant vector annihilated by all
\(\widehat K_\mu\) is a singular vector and generates a submodule.  In a
reflection-positive theory the radial inner product has a radical consisting
of null states, invariant under the conformal algebra because
\(\widehat P_\mu^\dagger=\widehat K_\mu\).  The cyclic physical multiplet
generated by the primary is obtained by quotienting \(M(\Delta,\rho)\) by this
null radical.  When the radical is the maximal proper submodule, this quotient
is the irreducible highest-weight module in the algebraic category.
\end{proposition}

\begin{proof}
The radial decomposition of the conformal algebra has
\([\widehat D_{\rm rad},\widehat P_\mu]=\widehat P_\mu\) and
\([\widehat D_{\rm rad},\widehat K_\mu]=-\widehat K_\mu\).  A primary state is
therefore a lowest-energy vector, and every algebraic descendant obtained from
it by the enveloping algebra may be represented, by the
Poincare--Birkhoff--Witt theorem, with all \(\widehat P_\mu\)'s to the left
of the primary vector.  Since the translations commute, the level-\(n\)
space is the symmetric tensor space in
\eqref{eq:conformal-verma-levels}.  A descendant annihilated by all
\(\widehat K_\mu\) satisfies the same lowest-energy condition as a primary and
therefore generates an invariant submodule.  In a reflection-positive theory
\(\widehat P_\mu^\dagger=\widehat K_\mu\); hence the radical of the radial
Gram form is invariant under both raising and lowering operators and may be
quotiented.  The quotient is the cyclic physical descendant module.  It is
irreducible once the radical has been identified with the maximal proper
submodule, equivalently once no further invariant descendant subspace remains.
\end{proof}

The determinant of the radial Gram form restricted to \(M(\Delta,\rho)_n\),
after decomposing into irreducible \(SO(D)\) summands, is the finite
conformal-algebra analogue of the Kac determinant.  Its zeros as a function
of \(\Delta\) locate possible singular descendants.  BGG embeddings and
resolutions organize the pattern of nested singular submodules.  Chapter
\ref{ch:volume-iii-unitarity-bounds-short-multiplets} computes the low-level
Gram factors needed for the unitarity bounds directly from the commutation
relations, without invoking the full determinant formula.

\begin{proposition}[Translation from the origin and descendant fields]
\label{prop:primary-translation-from-origin}
An operator at a general point is obtained by translating the insertion from
the origin, equivalently by its Taylor expansion in descendants:
\begin{equation}
  \mathcal O_a(x)
  =
  \sum_{k=0}^\infty
  \frac1{k!}
  x^{\mu_1}\cdots x^{\mu_k}
  \partial_{\mu_1}\cdots\partial_{\mu_k}\mathcal O_a(0).
  \label{eq:operator-from-origin-translation}
\end{equation}
In the Hermitian translation convention
\([\widehat P_\mu,\mathcal O(x)]=\ii\partial_\mu\mathcal O(x)\), the
corresponding state notation is
\[
  \ket{\mathcal O(x)}
  =
  \ee^{-\ii x^\mu\widehat P_\mu}\ket{\mathcal O}.
\]
\end{proposition}

\begin{proof}
The local field is an operator-valued distribution, so the displayed Taylor
series is a shorthand for equality after smearing with test functions in a
neighborhood on which the expansion converges as a radial state.  The
translation commutator gives
\[
  \frac{\dd}{\dd s}
  \left(
    \ee^{-\ii s x\cdot\widehat P}\mathcal O(0)
    \ee^{\ii s x\cdot\widehat P}
  \right)
  =
  x^\mu\partial_\mu\mathcal O(sx),
  \qquad
  \mathcal O(0)\ \text{at}\ s=0.
\]
Integrating this first-order equation gives the Taylor expansion
\eqref{eq:operator-from-origin-translation}.  Passing to the radial state
created by the insertion gives the state formula with the stated Hermitian
translation convention.
\end{proof}

\section{Spin Sources and Contact Terms}
\label{sec:primary-spin-sources-contact-terms}

\begin{definition}[Spin source convention and contact operator]
\label{def:primary-spin-source-contact-operator}
The primary module data must agree with the source-defined Ward identities.
Introduce a source \(\eta^a(x)\) valued in the dual representation
\(V_\rho^\vee\) and couple it by
\[
  \int\dd^D x\,\eta^a(x)\mathcal O_a(x).
\]
For a primary of dimension \(\Delta\), the flat-space source convention
associated with a conformal Killing vector \(\epsilon^\mu\) is
\begin{equation}
  \delta_\epsilon^{\rm src}\eta^a
  =
  \epsilon^\mu\partial_\mu\eta^a
  +
  (\Delta-D)\sigma_\epsilon\eta^a
  +
  \frac{\ii}{2}\omega_\epsilon^{\mu\nu}
  \eta^b(S_{\mu\nu})_b{}^a ,
  \label{eq:primary-source-transform}
\end{equation}
where
\[
  \sigma_\epsilon=\frac1D\partial_\mu\epsilon^\mu,
  \qquad
  \omega_{\epsilon,\mu\nu}
  =
  \frac12(\partial_\nu\epsilon_\mu-\partial_\mu\epsilon_\nu).
\]
The last term is the dual spin action.  The sign convention in
\eqref{eq:primary-source-transform} is the source convention of
Section~\ref{sec:cft-source-functionals-ward-contacts}; operator insertions
then acquire the contact operator
\begin{equation}
  \mathcal D_\epsilon\mathcal O_a(x)
  =
  -\epsilon^\mu(x)\partial_\mu\mathcal O_a(x)
  -\Delta\sigma_\epsilon(x)\mathcal O_a(x)
  -\frac{\ii}{2}\omega_\epsilon^{\mu\nu}(x)
  (S_{\mu\nu})_a{}^b\mathcal O_b(x).
  \label{eq:primary-contact-operator}
\end{equation}
\end{definition}

\begin{proposition}[Spinful local conformal-current Ward identity]
\label{prop:spinful-local-conformal-current-ward}
Differentiating the source Ward identity with respect to the sources and then
setting the sources to zero gives, for
\[
  X=\mathcal O_{1,a_1}(x_1)\cdots\mathcal O_{n,a_n}(x_n),
  \qquad x_i\ne x_j ,
\]
the local current Ward identity
\begin{equation}
  \partial_\mu
  \left\langle J^\mu_\epsilon(x)X\right\rangle
  =
  \sum_{i=1}^n
  \delta^{(D)}(x-x_i)
  \left\langle
    \mathcal O_{1,a_1}(x_1)\cdots
    \mathcal D_\epsilon^{(i)}\mathcal O_{i,a_i}(x_i)\cdots
    \mathcal O_{n,a_n}(x_n)
  \right\rangle
  +
  \sigma_\epsilon(x)\mathcal A(x;X).
  \label{eq:primary-spin-current-ward}
\end{equation}
Here \(\mathcal D_\epsilon^{(i)}\) uses the spin representation and scaling
dimension of the \(i\)th insertion.  Equation
\eqref{eq:primary-spin-current-ward} is the spinful version of
\eqref{eq:local-conformal-current-ward}.  With the small-sphere orientation
used in \eqref{eq:variation-by-small-charge},
\begin{equation}
  \ii\,Q_\epsilon(S_x)\mathcal O_a(x)
  =
  \mathcal D_\epsilon\mathcal O_a(x).
  \label{eq:small-sphere-primary-action}
\end{equation}
\end{proposition}

\begin{proof}
The source term is invariant to first order under the paired variations of
the source and the operator.  Integrating the transport term in
\eqref{eq:primary-source-transform} by parts gives the operator displacement
\(-\epsilon^\mu\partial_\mu\mathcal O_a\); the divergence part combines with
the source Weyl weight \(\Delta-D\) and the measure variation to give
\(-\Delta\sigma_\epsilon\mathcal O_a\); and the dual spin action on
\(\eta^a\) gives the spin action in
\eqref{eq:primary-contact-operator}.  Taking functional derivatives of the
source Ward identity with respect to the \(n\) sources produces a delta
function when the variation acts on the \(i\)th source, giving
\eqref{eq:primary-spin-current-ward}.  Integrating the current identity over
a small punctured ball around \(x\) and using the orientation convention of
Definition~\ref{def:small-surface-action-local-operator} gives
\eqref{eq:small-sphere-primary-action}.
\end{proof}

\section{Infinitesimal Action at a General Point}

\paragraph{Infinitesimal conformal action on a spin primary.}
Let
\[
  \epsilon^\mu(x)
  =
  a^\mu+\omega^\mu{}_\nu x^\nu+\lambda x^\mu
  +c^\mu x^2-2(c\cdot x)x^\mu
\]
be a conformal Killing vector in \(D\ge3\).  Define
\[
  \sigma_\epsilon(x)=\frac1D\partial_\mu\epsilon^\mu(x),
  \qquad
  \omega_{\epsilon,\mu\nu}(x)
  =
  \frac12(\partial_\nu\epsilon_\mu-\partial_\mu\epsilon_\nu).
\]
For a primary operator, the infinitesimal variation generated by the charge is
the contact operator \(\mathcal D_\epsilon\):
\begin{equation}
  \delta_\epsilon\mathcal O_a(x)
  =
  -\epsilon^\mu(x)\partial_\mu\mathcal O_a(x)
  -\Delta\sigma_\epsilon(x)\mathcal O_a(x)
  -\frac{\ii}{2}\omega_{\epsilon}^{\mu\nu}(x)
  (S_{\mu\nu})_a{}^b\mathcal O_b(x).
  \label{eq:primary-infinitesimal-transform}
\end{equation}
The first term moves the insertion point.  The second term is the local scale
weight.  The third term rotates the spin indices by the orthogonal part of the
Jacobian.
The formula is \eqref{eq:primary-contact-operator} written for a general
conformal Killing vector.  The special-conformal part is the only
nonconstant point dependence not visible from translations, rotations, and
dilatations, and it is checked explicitly below.

\begin{corollary}[Elementary generator actions]
\label{cor:primary-elementary-generator-actions}
For the four basic generators this gives
\begin{align}
  \delta_a\mathcal O(x)&=-a^\mu\partial_\mu\mathcal O(x),\\
  \delta_\lambda\mathcal O(x)&=-\lambda(x^\mu\partial_\mu+\Delta)\mathcal O(x),
  \label{eq:primary-translation-dilatation-transform}
\end{align}
and
\begin{align}
  \delta_\omega\mathcal O_a(x)
  &=
  -\omega^\mu{}_\nu x^\nu\partial_\mu\mathcal O_a(x)
  -\frac{\ii}{2}\omega^{\mu\nu}(S_{\mu\nu})_a{}^b\mathcal O_b(x),
  \\
  \delta_c\mathcal O_a(x)
  &=
  -\bigl(c^\mu x^2-2(c\cdot x)x^\mu\bigr)
  \partial_\mu\mathcal O_a(x)
  +2\Delta(c\cdot x)\mathcal O_a(x)
  \notag\\
  &\hspace{2.5em}
  -2\ii c^\mu x^\nu(S_{\mu\nu})_a{}^b\mathcal O_b(x).
  \label{eq:primary-sct-transform}
\end{align}
\end{corollary}

\begin{proof}
The special-conformal line is obtained directly from
\eqref{eq:primary-infinitesimal-transform}.  For
\[
  \epsilon_c^\mu(x)=c^\mu x^2-2(c\cdot x)x^\mu
\]
one has
\[
  \partial_\mu\epsilon_c^\mu=-2D(c\cdot x),
  \qquad
  \sigma_{\epsilon_c}=-2(c\cdot x),
\]
and
\[
  \omega_{\epsilon_c,\mu\nu}
  =
  \frac12(\partial_\nu\epsilon_{c,\mu}-\partial_\mu\epsilon_{c,\nu})
  =
  2(c_\mu x_\nu-c_\nu x_\mu).
\]
Substitution gives
\[
  -\Delta\sigma_{\epsilon_c}\mathcal O_a
  =
  2\Delta(c\cdot x)\mathcal O_a,
\]
and, using \(S_{\mu\nu}=-S_{\nu\mu}\),
\[
  -\frac{\ii}{2}\omega_{\epsilon_c}^{\mu\nu}
  (S_{\mu\nu})_a{}^b\mathcal O_b
  =
  -2\ii c^\mu x^\nu(S_{\mu\nu})_a{}^b\mathcal O_b.
\]
This proves \eqref{eq:primary-sct-transform} from the source-Ward contact
operator.  At \(x=0\), all three terms vanish for \(\epsilon_c\), so the same
formula is consistent with \(\widehat K_\mu\ket{\mathcal O_a}=0\) in the
state--operator correspondence.
\end{proof}

\section{Finite Transformation Law}

\begin{definition}[Conformal factor and local rotation]
\label{def:conformal-factor-local-rotation}
Let
\[
  f:U\longrightarrow f(U)
\]
be an orientation-preserving conformal map between open subsets of
\(\R^D\).  Its Jacobian is
\[
  M_f(x)^\mu{}_\nu=\frac{\partial f^\mu}{\partial x^\nu}.
\]
The conformal condition says that there is a positive function
\(\Omega_f(x)\) such that
\begin{equation}
  M_f(x)^T M_f(x)=\Omega_f(x)^2\mathbf 1 .
  \label{eq:jacobian-scale-rotation-factorization}
\end{equation}
The orthogonal matrix
\[
  R_f(x)=\Omega_f(x)^{-1}M_f(x)
\]
is the local rotation.  Since \(f\) is orientation preserving,
\(\Omega_f(x)^D=\det M_f(x)\).  The factors obey the cocycle identities
\[
  \Omega_{f\circ h}(x)=\Omega_f(h(x))\Omega_h(x),
  \qquad
  R_{f\circ h}(x)=R_f(h(x))R_h(x).
\]
\end{definition}

\begin{theorem}[Finite transformation law for a primary operator]
\label{thm:finite-primary-transformation-law}
The finite primary transformation law is the integrated form of
\eqref{eq:primary-infinitesimal-transform} on the connected conformal group:
\begin{equation}
  \mathcal O'_a(f(x))
  =
  \Omega_f(x)^{-\Delta}
  \rho(R_f(x))_a{}^b\mathcal O_b(x).
  \label{eq:finite-primary-transform}
\end{equation}
\end{theorem}

\begin{proof}
Let \(f_s\) be the one-parameter flow of a
conformal Killing vector,
\[
  \frac{\dd}{\dd s}f_s^\mu(x)=\epsilon^\mu(f_s(x)),
  \qquad
  f_0(x)=x,
\]
and write \(M_s=\partial f_s/\partial x=\Omega_s R_s\).  At \(s=0\),
\begin{equation}
  \left.\frac{\dd}{\dd s}\Omega_s(x)\right|_{s=0}
  =
  \sigma_\epsilon(x),
  \qquad
  \left.\frac{\dd}{\dd s}R_s(x)_\mu{}^\nu\right|_{s=0}
  =
  \omega_{\epsilon,\mu}{}^\nu(x),
  \label{eq:primary-finite-infinitesimal-jacobian}
\end{equation}
because the conformal Killing equation decomposes
\(\partial_\nu\epsilon_\mu\) into its trace
\(\sigma_\epsilon\delta_{\mu\nu}\) and antisymmetric part
\(\omega_{\epsilon,\mu\nu}\).  The transformed field at a fixed final point
\(y\) is obtained from \(x=f_s^{-1}(y)\):
\[
  \mathcal O_{s,a}'(y)
  =
  \Omega_s(f_s^{-1}y)^{-\Delta}
  \rho(R_s(f_s^{-1}y))_a{}^b
  \mathcal O_b(f_s^{-1}y).
\]
Since \(\left.\frac{\dd}{\dd s}f_s^{-1}(y)\right|_{s=0}
=-\epsilon(y)\) and
\[
  \left.\frac{\dd}{\dd s}\rho(R_s)_a{}^b\right|_{s=0}
  =
  -\frac{\ii}{2}\omega_\epsilon^{\mu\nu}
  (S_{\mu\nu})_a{}^b ,
\]
differentiation at \(s=0\) gives exactly
\eqref{eq:primary-infinitesimal-transform}.  Conversely, the vector fields
close under the Lie bracket and the factors \(\Omega_f,R_f\) obey the cocycle
identities above; hence the finite law is the integrated representation of the
infinitesimal contact action on every connected patch where the conformal map
is defined.
\end{proof}

\begin{corollary}[Elementary finite conformal transformations]
\label{cor:elementary-finite-primary-transformations}
The elementary generators make the signs transparent.  A translation
\(f_a(x)=x+a\) has \(\Omega=1\), \(R=1\), and gives
\(\delta_a\mathcal O=-a^\mu\partial_\mu\mathcal O\) at fixed final coordinate.
A dilatation \(f_\lambda(x)=\ee^\lambda x\) has
\(\Omega=\ee^\lambda\), \(R=1\), and therefore
\[
  \mathcal O'(\ee^\lambda x)=\ee^{-\Delta\lambda}\mathcal O(x),
  \qquad
  \mathcal O'(y)=\ee^{-\Delta\lambda}\mathcal O(\ee^{-\lambda}y),
\]
whose derivative is
\(-\lambda(y^\mu\partial_\mu+\Delta)\mathcal O(y)\).  A rotation has
\(\Omega=1\) and \(R\) equal to its orthogonal matrix, so exponentiating
\(\rho(R)=\exp[-\frac{\ii}{2}\omega^{\mu\nu}S_{\mu\nu}]\) gives the spin term
in \eqref{eq:primary-infinitesimal-transform}.  Finally the finite special
conformal transformation is
\[
  f_c=I\circ T_c\circ I,
  \qquad
  f_c(x)=\frac{x+c\,x^2}{1+2c\cdot x+c^2x^2};
\]
expanding at \(c=0\) gives
\(\epsilon_c^\mu=c^\mu x^2-2(c\cdot x)x^\mu\), so the differentiated finite law
reproduces \eqref{eq:primary-sct-transform}.
\end{corollary}

\begin{remark}[Tensor primary convention]
For tensor representations this formula is the tensor transformation law with
the Jacobian decomposed into the scale factor \(\Omega_f(x)\) and the
orthogonal rotation \(R_f(x)\).  At coincident points the corresponding
statement is made at the level of the transformed source functional; the
displayed formula is its restriction to noncoincident insertions.
\end{remark}

\begin{proposition}[Finite transformation of tensor primaries]
\label{prop:finite-primary-tensor-transformations}
For a covariant rank-\(s\) primary tensor
\(\mathcal O_{\mu_1\cdots\mu_s}\), the same formula can be written without
abstract representation notation as
\begin{equation}
  \mathcal O'_{\mu_1\cdots\mu_s}(f(x))
  =
  \left(\det\frac{\partial f}{\partial x}\right)^{-(\Delta-s)/D}
  \frac{\partial x^{\nu_1}}{\partial f^{\mu_1}}\cdots
  \frac{\partial x^{\nu_s}}{\partial f^{\mu_s}}\,
  \mathcal O_{\nu_1\cdots\nu_s}(x).
  \label{eq:finite-primary-covariant-tensor}
\end{equation}
For a contravariant rank-\(s\) primary tensor,
\begin{equation}
  \mathcal O'^{\mu_1\cdots\mu_s}(f(x))
  =
  \left(\det\frac{\partial f}{\partial x}\right)^{-(\Delta+s)/D}
  \frac{\partial f^{\mu_1}}{\partial x^{\nu_1}}\cdots
  \frac{\partial f^{\mu_s}}{\partial x^{\nu_s}}\,
  \mathcal O^{\nu_1\cdots\nu_s}(x).
  \label{eq:finite-primary-contravariant-tensor}
\end{equation}
The different determinant powers in
\eqref{eq:finite-primary-covariant-tensor} and
\eqref{eq:finite-primary-contravariant-tensor} follow from
\(M_f=\Omega_f R_f\).  Each covariant index contributes one factor
\(\Omega_f^{-1}\) from \(M_f^{-1}\), each contravariant index contributes one
factor \(\Omega_f\) from \(M_f\), and the remaining determinant factor supplies
the primary weight \(\Omega_f^{-\Delta}\).
\end{proposition}

\begin{proof}
For a conformal diffeomorphism \(f\), the Jacobian matrix
\[
  M^\mu{}_\nu(x)=\frac{\partial f^\mu}{\partial x^\nu}
\]
satisfies \(M^T M=\Omega_f(x)^2\mathbf 1\).  Hence
\[
  R^\mu{}_\nu(x):=\Omega_f(x)^{-1}M^\mu{}_\nu(x)
\]
is orthogonal on the Euclidean tangent space, and
\[
  \det M=\Omega_f^D\det R.
\]
On the connected orientation-preserving component, \(\det R=1\), so
\(\Omega_f=(\det M)^{1/D}\); on an orientation-reversing component the
orientation sign is part of the \(O(D)\) tensor representation and the
positive scale is \(\Omega_f=|\det M|^{1/D}\).  The formulas above are written
for the orientation-preserving component, which is the component used in the
local conformal Ward identities.

The abstract primary law proved in
Theorem~\ref{thm:finite-primary-transformation-law} is
\[
  \mathcal O'_a(f(x))
  =
  \Omega_f(x)^{-\Delta}\rho(R_f(x))_a{}^b\mathcal O_b(x).
\]
For a covariant vector index, the orthogonal representation acts by
\[
  (R^{-1})^\nu{}_\mu
  =
  \Omega_f\frac{\partial x^\nu}{\partial f^\mu},
\]
because \(M^{-1}=\partial x/\partial f\).  For a contravariant vector index it
acts by
\[
  R^\mu{}_\nu
  =
  \Omega_f^{-1}\frac{\partial f^\mu}{\partial x^\nu}.
\]
Tensor-product representations multiply these factors index by index.  Thus a
covariant rank-\(s\) tensor acquires
\[
  \Omega_f^{-\Delta}\Omega_f^s
  \frac{\partial x^{\nu_1}}{\partial f^{\mu_1}}\cdots
  \frac{\partial x^{\nu_s}}{\partial f^{\mu_s}}
  =
  (\det M)^{-(\Delta-s)/D}
  \frac{\partial x^{\nu_1}}{\partial f^{\mu_1}}\cdots
  \frac{\partial x^{\nu_s}}{\partial f^{\mu_s}},
\]
which is \eqref{eq:finite-primary-covariant-tensor}.  Similarly a
contravariant rank-\(s\) tensor acquires
\[
  \Omega_f^{-\Delta}\Omega_f^{-s}
  \frac{\partial f^{\mu_1}}{\partial x^{\nu_1}}\cdots
  \frac{\partial f^{\mu_s}}{\partial x^{\nu_s}}
  =
  (\det M)^{-(\Delta+s)/D}
  \frac{\partial f^{\mu_1}}{\partial x^{\nu_1}}\cdots
  \frac{\partial f^{\mu_s}}{\partial x^{\nu_s}},
\]
which is \eqref{eq:finite-primary-contravariant-tensor}.
\end{proof}

\begin{remark}[Finite covariance of primary correlators]
\label{rem:primary-correlator-finite-covariance}
If the vacuum is invariant under the conformal transformation, correlation
functions of primaries obey
\begin{equation}
  \left\langle
    \prod_{i=1}^n\mathcal O_{i,a_i}(f(x_i))
  \right\rangle
  =
  \prod_{i=1}^n
  \Omega_f(x_i)^{-\Delta_i}
  \rho_i(R_f(x_i))_{a_i}{}^{b_i}
  \left\langle
    \prod_{i=1}^n\mathcal O_{i,b_i}(x_i)
  \right\rangle .
  \label{eq:primary-correlator-covariance}
\end{equation}
Equation \eqref{eq:primary-correlator-covariance} is the finite form of the
global Ward identities on the configuration space of distinct insertion
points, with contact extensions fixed by the same source convention as in
\eqref{eq:primary-spin-current-ward}.
Let \(U_f\) denote the Hilbert-space or Euclidean functional action of the
connected conformal transformation on the common domain of separated
insertions.  Vacuum invariance means that inserting
\(U_f^{-1}U_f\) in the expectation value does not change it:
\[
  \left\langle\prod_i\mathcal O_{i,a_i}(f(x_i))\right\rangle
  =
  \left\langle
  U_f^{-1}\left(\prod_i\mathcal O_{i,a_i}(f(x_i))\right)U_f
  \right\rangle .
\]
On the configuration space \(x_i\ne x_j\), the conjugated product is the
ordinary product of the individually transformed fields, because no two
insertions collide while the transformation is applied.  Substituting
\eqref{eq:finite-primary-transform} for each primary therefore produces one
factor \(\Omega_f(x_i)^{-\Delta_i}\rho_i(R_f(x_i))\) for each insertion and
gives \eqref{eq:primary-correlator-covariance}.  Extensions to collision
diagonals are not determined by this separated-point argument; they are fixed
by the source Ward identity \eqref{eq:primary-spin-current-ward}.
\end{remark}

\section{Inversion}

\paragraph{Inversion of scalar and spin primaries.}
The inversion map
\[
  I(x)^\mu=\frac{x^\mu}{x^2}
\]
has
\[
  \Omega_I(x)=\frac1{x^2},
  \qquad
  R_I(x)^\mu{}_\nu
  =
  \delta^\mu{}_\nu-2\frac{x^\mu x_\nu}{x^2}.
\]
For a scalar primary,
\begin{equation}
  \mathcal O'(I(x))
  =
  |x|^{2\Delta}\mathcal O(x).
  \label{eq:scalar-primary-inversion}
\end{equation}
This formula, together with translations, rotations, and dilatations,
determines the separated-point covariance constraints used below.  The matrix
\(R_I(x)\) lies in \(O(D)\) and has determinant \(-1\).  For operators with
spin, inversion therefore requires a specified extension of the spin action
from \(SO(D)\) to the relevant disconnected orthogonal or pin action; once
that choice is made, \(R_I(x)\) acts on each spin index through that
representation.

Differentiating \(I^\mu=x^\mu/x^2\) gives
\[
  \frac{\partial I^\mu}{\partial x^\nu}
  =
  \frac1{x^2}
  \left(
    \delta^\mu{}_\nu-2\frac{x^\mu x_\nu}{x^2}
  \right).
\]
The matrix in parentheses squares to the identity and has determinant
\(-1\), since it reflects the component parallel to \(x^\mu\) and fixes the
orthogonal hyperplane.  Thus \(\Omega_I=1/x^2\) and
\(R_I=\delta-2xx/x^2\).  Substituting these factors into
\eqref{eq:finite-primary-transform} gives
\eqref{eq:scalar-primary-inversion} for scalar primaries and the stated
orthogonal or pin action for spin primaries.

\section{Descendants}

\paragraph{Descendant transformation laws from primary data.}
The descendants of a primary are obtained by applying derivatives.  Their
transformation laws are obtained by differentiating
\eqref{eq:finite-primary-transform} and by decomposing the result into
irreducible \(SO(D)\) representations.  Once the primary dimension, spin
representation, and normalization are fixed, the action of the whole conformal
algebra on the generated multiplet is fixed.  Concretely, every descendant at
level \(n\) is represented algebraically by an element of
\(\operatorname{Sym}^n(\mathbb C^D)\otimes V_\rho\) in
\eqref{eq:conformal-verma-levels}, modulo any null or equation-of-motion
relations.  The conformal generators act on this space by the commutation
relations of \(\mathfrak{so}(D+2,\mathbb C)\).  Equivalently, at separated
points, differentiating \eqref{eq:finite-primary-transform} with respect to
the insertion coordinate gives the finite transformation of derivatives of
\(\mathcal O\).  Decomposing the resulting tensor products into irreducible
\(SO(D)\) summands changes the basis of descendants while preserving the
representation generated by the primary datum.
