\chapter{Conformal Charges and Ward Identities}
\label{ch:volume-iii-conformal-charges-ward-identities}
\ConventionsEuclideanCFT


\section{Charges as Surface Operators}

\begin{definition}[Conformal surface charge]
\label{def:conformal-surface-charge}
Let \(\epsilon^\mu(x)\partial_\mu\) be a conformal Killing vector on flat
Euclidean space, and let
\[
  J^\mu_\epsilon(x)=T^\mu{}_\nu(x)\epsilon^\nu(x)
\]
be the corresponding conserved current at separated points.  If
\(\Sigma\subset\R^D\) is an oriented smooth codimension-one surface with unit
normal \(n_\mu\), define the charge supported on \(\Sigma\) by
\begin{equation}
  Q_\epsilon(\Sigma)
  =
  \int_\Sigma \dd\sigma\,n_\mu T^\mu{}_\nu(x)\epsilon^\nu(x).
  \label{eq:cft-charge-surface}
\end{equation}
Here \(\dd\sigma\) is the induced volume element.
\end{definition}

\paragraph{Surface deformation of a conserved charge.}
Let \(\Sigma_0\) and \(\Sigma_1\) be homologous oriented smooth
codimension-one surfaces in a region \(U\subset\R^D\) that contains no operator
insertions and no boundary.  If
\(\partial_\mu J^\mu_\epsilon=0\) as a separated-point distribution on
\(U\), then \(Q_\epsilon(\Sigma_0)=Q_\epsilon(\Sigma_1)\) inside every
correlation function whose other insertions lie outside the deformation
region.

Let \(V\subset U\) be the oriented region swept out by the deformation, with
\(\partial V=\Sigma_1-\Sigma_0\).  Smearing the distributional conservation
law with a test function that is one on \(V\) and whose support avoids all
insertions gives
\[
  0
  =
  \int_V\dd^D x\,\partial_\mu J^\mu_\epsilon(x)
  =
  \int_{\Sigma_1}\dd\sigma\,n_\mu J^\mu_\epsilon
  -
  \int_{\Sigma_0}\dd\sigma\,n_\mu J^\mu_\epsilon .
\]
The last equality is Stokes' theorem for the smoothed distributional pairing;
shrinking the transition region of the test function to zero gives the
surface statement.  No contact term appears because the support of the
deformation avoids the insertion diagonals.

This deformation property is the local content of a conserved charge.  A
charge may be evaluated on a large sphere surrounding several insertions, or
on a union of small spheres surrounding them separately.  The equality of
these two evaluations is the Ward identity.

\begin{definition}[Lorentzian charge commutator across an insertion]
\label{def:lorentzian-charge-commutator-across-insertion}
The Lorentzian operator meaning of the same construction is the following.
Let \(\widehat Q_\epsilon(\Sigma)\) be the flux of the conserved current
through a spacelike Cauchy surface \(\Sigma\), with future-directed unit normal.
If a local operator \(\widehat{\mathcal O}(0)\) is inserted between two nearby
Cauchy surfaces \(\Sigma_-\) and \(\Sigma_+\), then
\begin{equation}
  [\widehat Q_\epsilon,\widehat{\mathcal O}(0)]
  =
  \widehat Q_\epsilon(\Sigma_+)\widehat{\mathcal O}(0)
  -
  \widehat{\mathcal O}(0)\widehat Q_\epsilon(\Sigma_-).
  \label{eq:lorentzian-charge-commutator-by-slices}
\end{equation}
Here conservation identifies the charge on \(\Sigma_\pm\) with the same
abstract symmetry generator when no insertion lies between the two surfaces.
Equation \eqref{eq:lorentzian-charge-commutator-by-slices} is the definition
of the commutator after the charge has been represented on the Hilbert spaces
immediately to the future and past of the insertion.  After Euclidean
continuation, the two surfaces close into a small hypersurface surrounding the
insertion, and the commutator becomes the small-surface action used below.
\end{definition}

Figure~\ref{fig:charge-commutator-slices} records this Lorentzian-to-Euclidean
translation of a charge commutator: the difference between future and past
charge slices becomes a closed surface surrounding the insertion.

\begin{figure}[H]
  \centering
  \resizebox{0.88\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      slice/.style={green!45!black,line width=0.95pt},
      insertion/.style={circle,fill=violet!80!black,inner sep=1.8pt},
      arrow/.style={-{Latex[length=2mm]},line width=0.8pt}
    }
    \draw[arrow] (-4.7,-1.8)--(-4.7,2.25) node[above] {time};
    \draw[arrow] (-5.0,-1.45)--(3.9,-1.45) node[right] {space};
    \draw[slice]
      (-4.1,-0.65) .. controls (-2.8,-0.35) and (-1.45,-0.85) ..
      (0.0,-0.50) .. controls (1.45,-0.15) and (2.35,-0.55) .. (3.25,-0.25);
    \draw[slice]
      (-4.1,0.95) .. controls (-2.8,1.25) and (-1.45,0.75) ..
      (0.0,1.10) .. controls (1.45,1.45) and (2.35,1.05) .. (3.25,1.35);
    \node[green!45!black] at (3.55,-0.25) {\(\Sigma_-\)};
    \node[green!45!black] at (3.55,1.35) {\(\Sigma_+\)};
    \node[insertion] at (-0.35,0.25) {};
    \node[violet!80!black,below right] at (-0.25,0.19) {\(\widehat{\mathcal O}(0)\)};
    \draw[blue!65!black,dashed,line width=0.8pt]
      (-1.35,-0.78) .. controls (-1.65,-0.05) and (-1.25,0.78) ..
      (-0.42,1.15) .. controls (0.55,1.58) and (1.35,0.75) ..
      (1.22,-0.05) .. controls (1.10,-0.88) and (-0.35,-1.10) ..
      (-1.35,-0.78);
    \node[blue!65!black,align=center] at (1.95,0.35)
      {closed surface\\after continuation};
  \end{tikzpicture}
  }
  \caption{A Lorentzian commutator with a conserved charge is represented by
  the difference between charge insertions on slices immediately after and
  before the local operator.  In Euclidean signature this difference is encoded
  by a closed surface enclosing the insertion.}
  \label{fig:charge-commutator-slices}
\end{figure}

Figure~\ref{fig:ward-surface-deformation} displays the corresponding Euclidean
Ward-identity deformation: a closed charge surface can be moved through regions
free of insertions, and crossing an insertion gives the local symmetry action.

\begin{figure}[H]
  \centering
  \resizebox{0.95\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      insertion/.style={circle,fill=violet!80!black,inner sep=1.8pt},
      surface/.style={green!45!black,line width=0.85pt},
      arrow/.style={-{Latex[length=2mm]},line width=0.8pt}
    }
    \draw[thick] (-5.4,-2.05) rectangle (5.4,2.15);
    \node[anchor=west] at (-5.25,2.45) {\(\R^D\)};

    \draw[surface]
      (-4.35,-1.25) .. controls (-4.85,0.10) and (-3.85,1.85) .. (-1.95,1.75)
      .. controls (-0.85,2.15) and (0.85,2.05) .. (2.05,1.50)
      .. controls (4.05,1.10) and (4.65,-0.65) .. (3.55,-1.40)
      .. controls (2.45,-2.10) and (0.35,-1.65) .. (-1.20,-1.75)
      .. controls (-2.80,-1.85) and (-4.00,-1.80) .. (-4.35,-1.25);
    \node[green!45!black] at (3.95,1.65) {\(\Sigma\)};

    \foreach \x/\y/\lab/\slabel in {
      -2.65/0.55/{x_1}/{S_1},
       0.05/0.85/{x_2}/{S_2},
       2.45/-0.55/{x_n}/{S_n}
    }{
      \node[insertion] at (\x,\y) {};
      \node[violet!80!black,above right] at (\x,\y) {\(\mathcal O(\lab)\)};
      \draw[blue!65!black,line width=0.75pt] (\x,\y) circle (0.55);
      \node[blue!65!black] at (\x+0.52,\y+0.62) {\(\slabel\)};
    }
    \node at (1.35,0.15) {\(\cdots\)};

    \draw[arrow] (5.85,0.55)--(7.15,0.55);
    \node[align=center] at (6.50,1.05) {deform\\surface};

    \begin{scope}[shift={(9.55,0)}]
      \draw[thick] (-2.05,-2.05) rectangle (2.05,2.15);
      \foreach \x/\y/\lab/\slabel in {
        -1.05/0.55/{x_1}/{S_1},
         0.20/0.85/{x_2}/{S_2},
         1.00/-0.55/{x_n}/{S_n}
      }{
        \node[insertion] at (\x,\y) {};
        \node[violet!80!black,above right] at (\x,\y) {\(\mathcal O(\lab)\)};
        \draw[surface] (\x,\y) circle (0.43);
        \node[green!45!black] at (\x+0.40,\y+0.50) {\(\slabel\)};
      }
      \node at (0.68,0.10) {\(\cdots\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{A conserved conformal charge can be moved from a large enclosing
  surface to small enclosing surfaces around the insertions.  The Ward identity
  is this equality inside correlation functions.}
  \label{fig:ward-surface-deformation}
\end{figure}

\section{Operator Variations from Small Spheres}

\begin{definition}[Small-surface action on a local operator]
\label{def:small-surface-action-local-operator}
Let \(S_x\) be a small sphere enclosing a single insertion at \(x\), with
orientation inherited from the boundary of a small ball removed from the
correlation-function domain.  The infinitesimal action of the symmetry on the
operator \(\mathcal O(x)\) is defined by
\begin{equation}
  \delta_\epsilon\mathcal O(x)
  =
  \ii\,Q_\epsilon(S_x)\mathcal O(x)
  \label{eq:variation-by-small-charge}
\end{equation}
in Euclidean radial-ordering conventions.  In Lorentzian operator language,
this convention corresponds to
\[
  \delta_\epsilon\widehat{\mathcal O}
  =
  \ii[\widehat Q_\epsilon,\widehat{\mathcal O}].
\]
Changing the orientation convention changes both sides of
\eqref{eq:variation-by-small-charge} by the same sign; the Ward identity
below is then unchanged.
\end{definition}

\begin{hypothesis}[Vacuum and large-boundary behavior]
\label{hyp:conformal-charge-vacuum-boundary}
For the conformal Killing vector \(\epsilon^\mu\partial_\mu\) and the
correlators under consideration, assume the following two properties.
\begin{enumerate}[label=(\roman*)]
  \item The vacuum is invariant under the corresponding charge.  In Lorentzian
  notation this means
  \[
    \widehat Q_\epsilon\ket{\vac}=0,
    \qquad
    \bra{\vac}\widehat Q_\epsilon=0
  \]
  on the common invariant domain on which the correlators are defined.
  \item If \(\Sigma_R\) is a family of smooth enclosing surfaces whose distance
  from the insertions tends to infinity, then clustering makes the large
  boundary flux vanish:
  \[
    \lim_{R\to\infty}
    \left\langle
      Q_\epsilon(\Sigma_R)
      \prod_{i=1}^n\mathcal O_i(x_i)
    \right\rangle
    =
    0 .
  \]
  More explicitly, with
  \[
    Q_\epsilon(\Sigma_R)
    =
    \int_{\Sigma_R}\dd S_\mu\,
    \epsilon_\nu(x)T^{\mu\nu}(x),
  \]
  the connected correlator
  \[
    \left\langle
      T^{\mu\nu}(x)\prod_{i=1}^n\mathcal O_i(x_i)
    \right\rangle_{\!c}
  \]
  must decay at large \(|x|\) fast enough to overcome the area growth of
  \(\Sigma_R\) and the polynomial growth of the displayed conformal Killing
  vectors.  This is the cluster-decomposition input in the deformation of the
  large surface to the small spheres \(S_i\).
\end{enumerate}
\end{hypothesis}

\paragraph{Global conformal Ward identity from surface deformation.}
Under Hypothesis~\ref{hyp:conformal-charge-vacuum-boundary}, for separated
insertions,
\begin{equation}
  0
  =
  \left\langle
    Q_\epsilon(\Sigma)
    \prod_{i=1}^n\mathcal O_i(x_i)
  \right\rangle
  =
  \sum_{i=1}^n
  \left\langle
    \mathcal O_1(x_1)\cdots
    Q_\epsilon(S_i)\mathcal O_i(x_i)\cdots
    \mathcal O_n(x_n)
  \right\rangle ,
  \label{eq:surface-ward-decomposition}
\end{equation}
where \(\Sigma\) encloses all insertions and \(S_i\) is a small sphere
enclosing \(x_i\).  Using
\eqref{eq:variation-by-small-charge}, this becomes
\begin{equation}
  \sum_{i=1}^n
  \left\langle
    \mathcal O_1(x_1)\cdots
    \delta_\epsilon\mathcal O_i(x_i)\cdots
    \mathcal O_n(x_n)
  \right\rangle
  =
  0.
  \label{eq:global-conformal-ward-identity}
\end{equation}
This is the global Ward identity associated with the conformal Killing vector
\(\epsilon^\mu\).  It is a statement about distributions on the configuration
space of distinct points, completed by contact terms on diagonals when
operator products collide.

By vacuum invariance, the charge on a sufficiently large surface enclosing all
insertions has zero expectation value after the large-boundary limit in
Hypothesis~\ref{hyp:conformal-charge-vacuum-boundary}.  By
the surface-deformation argument at the beginning of the chapter, that large
surface can be deformed through the punctured configuration space to the union
of small surfaces \(S_i\) surrounding the insertions.  The induced orientation
on the small boundaries is the one used in
Definition~\ref{def:small-surface-action-local-operator}, and hence each
small surface gives
\[
  Q_\epsilon(S_i)\mathcal O_i(x_i)
  =
  -\ii\,\delta_\epsilon\mathcal O_i(x_i).
\]
Multiplying the equality of surface insertions by \(\ii\) gives
\eqref{eq:global-conformal-ward-identity}.  This derivation is carried out on the
configuration space of distinct points; the extension to coincident
configurations is a further source-convention statement about contact terms.

\section{Source-Derived Local Ward Identities}

The surface identity is the integral form of the source Ward identities
introduced in Section~\ref{sec:cft-source-functionals-ward-contacts}.  Let
\[
  X=\mathcal O_1(x_1)\cdots\mathcal O_n(x_n),
  \qquad x_i\ne x_j ,
\]
where the \(\mathcal O_i\) are scalar local operators.  In the source chart
used for the stress tensor and the operators, the following distributional
identities are the local form of the conformal Ward identity.

\begin{proposition}[Local conformal-current Ward identity]
\label{prop:local-conformal-current-ward}
Assume that the stress tensor is symmetric and that scalar primary sources are
defined in the same source chart as the metric-source stress tensor.  Let
\(\mathcal O_i\) have scaling dimensions \(\Delta_i\), and set
\[
  \sigma_\epsilon=\frac1D\partial_\rho\epsilon^\rho,
  \qquad
  J^\mu_\epsilon=T^\mu{}_\nu\epsilon^\nu .
\]
The translation Ward identity is
\begin{equation}
  \partial_\mu
  \left\langle
    T^{\mu\nu}(x)X
  \right\rangle
  =
  -\sum_{i=1}^n
  \delta^{(D)}(x-x_i)
  \left\langle
    \mathcal O_1(x_1)\cdots
    \partial_i^\nu\mathcal O_i(x_i)\cdots
    \mathcal O_n(x_n)
  \right\rangle .
  \label{eq:local-translation-ward}
\end{equation}
Here \(\partial_i^\nu\) differentiates the \(i\)th insertion coordinate.  A
spin-carrying insertion adds the local representation term dictated by the
diffeomorphism action on its source.  Equation
\eqref{eq:local-translation-ward} is equivalent to the smeared identity
\eqref{eq:cft-smeared-translation-ward}; the displayed delta functions record
the extension of the separated-point distribution across the diagonals
\(x=x_i\).

The flat-space trace Ward identity in the same source chart is
\begin{equation}
  \left\langle
    T^\mu{}_\mu(x)X
  \right\rangle
  =
  -\sum_{i=1}^n
  \Delta_i\,\delta^{(D)}(x-x_i)
  \left\langle X\right\rangle
  +
  \mathcal A(x;X).
  \label{eq:local-trace-ward}
\end{equation}
The distribution \(\mathcal A(x;X)\) is the local anomaly or contact term
obtained by differentiating the local source functional
\(\mathcal A_\sigma[X]\) in \eqref{eq:cft-smeared-weyl-ward} with respect to
the Weyl parameter \(\sigma(x)\).  It vanishes on the configuration space of
distinct flat-space points; on collision diagonals or curved backgrounds it is
part of the specified source convention.

If \(\epsilon^\mu\) is a conformal Killing vector, then
\begin{equation}
  \partial_\mu
  \left\langle J^\mu_\epsilon(x)X\right\rangle
  =
  \sum_{i=1}^n
  \delta^{(D)}(x-x_i)
  \left\langle
    \mathcal O_1(x_1)\cdots
    \delta_\epsilon\mathcal O_i(x_i)\cdots
    \mathcal O_n(x_n)
  \right\rangle
  +
  \sigma_\epsilon(x)\mathcal A(x;X),
  \label{eq:local-conformal-current-ward}
\end{equation}
where for a scalar primary
\[
  \delta_\epsilon\mathcal O_i(x_i)
  =
  -\epsilon^\nu(x_i)\partial_{i,\nu}\mathcal O_i(x_i)
  -\Delta_i\sigma_\epsilon(x_i)\mathcal O_i(x_i).
\]
For spin \(\rho_i\), \(\delta_\epsilon\mathcal O_i\) includes the local
rotation term displayed later in
\eqref{eq:primary-infinitesimal-transform}.
\end{proposition}

\begin{proof}
The translation identity is the functional derivative of the metric-source
diffeomorphism Ward identity with respect to the scalar sources, followed by
setting all sources to their flat background values.  The sign is fixed by the
active variation
\(\delta_\epsilon\mathcal O_i=-\epsilon^\nu\partial_{i,\nu}\mathcal O_i\)
for translations.  The trace identity is the corresponding functional
derivative of the Weyl Ward identity in the same source chart; the displayed
\(\Delta_i\)-term comes from the Weyl weight of the scalar source, while
\(\mathcal A(x;X)\) is the differentiated local Weyl anomaly functional.

For a conformal Killing vector,
\[
  \partial_\mu\epsilon_\nu+\partial_\nu\epsilon_\mu
  =
  2\sigma_\epsilon\delta_{\mu\nu}.
\]
Using symmetry of \(T^{\mu\nu}\),
\[
  \partial_\mu\langle J^\mu_\epsilon(x)X\rangle
  =
  \epsilon_\nu(x)\partial_\mu
  \langle T^{\mu\nu}(x)X\rangle
  +
  \sigma_\epsilon(x)
  \langle T^\mu{}_\mu(x)X\rangle .
\]
Substituting \eqref{eq:local-translation-ward} and
\eqref{eq:local-trace-ward} gives
\eqref{eq:local-conformal-current-ward}, with the delta functions evaluating
\(\epsilon^\mu\) and \(\sigma_\epsilon\) at the insertion points.
\end{proof}

Integrating
\eqref{eq:local-conformal-current-ward} over a region whose boundary is
\(\Sigma\cup_i(-S_i)\) gives the surface decomposition
\eqref{eq:surface-ward-decomposition}; if the large boundary contribution
vanishes and the anomaly term integrates to zero in the chosen background, the
global Ward identity \eqref{eq:global-conformal-ward-identity} follows.

\section{Conformal Charges on the Cylinder}

\begin{proposition}[Cylinder conformal charges and radial adjointness]
\label{prop:cylinder-conformal-charges-radial-adjointness}
Under the Weyl map
\[
  \R^D\setminus\{0\}\longrightarrow \R_\tau\times S^{D-1},
  \qquad
  x^\mu=\ee^\tau n^\mu,
  \qquad
  n^\mu n_\mu=1,
\]
closed spheres surrounding the origin become constant-\(\tau\) slices.  The
dilatation charge becomes the cylinder Hamiltonian, rotations become the
rotations of \(S^{D-1}\), and the translation and special conformal charges
become mutually adjoint raising and lowering operators:
\begin{equation}
  \widehat P_\mu^\dagger=\widehat K_\mu .
  \label{eq:p-k-adjoint-cylinder}
\end{equation}
This adjointness is a consequence of reflection positivity under
\(\tau\mapsto-\tau\), which is inversion in flat space.  It is the positivity
input used below to derive unitarity bounds.
\end{proposition}

\begin{proof}
The map \(x^\mu=\ee^\tau n^\mu\) sends the flat metric on
\(\R^D\setminus\{0\}\) to
\(\dd s^2=\ee^{2\tau}(\dd\tau^2+\dd\Omega_{D-1}^2)\); after the Weyl
rescaling by \(\ee^{-2\tau}\), spheres \(|x|=\ee^\tau\) are the constant
cylinder-time slices.  The conformal Killing vector \(x^\mu\partial_\mu\)
therefore becomes \(\partial_\tau\), so its charge is the generator of
Euclidean cylinder translations.  Reflection positivity pairs an insertion at
\(\tau\) with its reflected insertion at \(-\tau\), and reflection in flat
coordinates is inversion \(x^\mu\mapsto x^\mu/x^2\).  Inversion conjugates
translations to special conformal transformations; hence the positive radial
inner product identifies the adjoint of the translation charge with the
special conformal charge, giving \eqref{eq:p-k-adjoint-cylinder}.  Rotations
commute with the Weyl factor and act as the ordinary isometries of
\(S^{D-1}\).
\end{proof}

More explicitly, let
\[
  \Pi_{\mu\nu}(n)=\delta_{\mu\nu}-n_\mu n_\nu
\]
be the projector onto the tangent space of \(S^{D-1}\) at \(n\).  Acting on
functions on \(\R_\tau\times S^{D-1}\), the flat-space conformal Killing
vectors become
\begin{align}
  D_{\rm rad}
  &=
  \partial_\tau,
  \\
  P_\mu
  &=
  \ee^{-\tau}\left(n_\mu\partial_\tau
  +\Pi_{\mu\nu}(n)\frac{\partial}{\partial n_\nu}\right),
  \\
  K_\mu
  &=
  \ee^\tau\left(-n_\mu\partial_\tau
  +\Pi_{\mu\nu}(n)\frac{\partial}{\partial n_\nu}\right),
  \label{eq:pk-cylinder-vector-fields}
\end{align}
with rotations acting as vector fields tangent to the sphere.  Thus inversion
\(x^\mu\mapsto x^\mu/x^2\), which is \(\tau\mapsto-\tau\), exchanges
\(P_\mu\) and \(K_\mu\).  As vector fields their weights under
\(D_{\rm rad}^{\rm vec}=\partial_\tau\) are
\[
  [\partial_\tau,P_\mu^{\rm vec}]=-P_\mu^{\rm vec},
  \qquad
  [\partial_\tau,K_\mu^{\rm vec}]=K_\mu^{\rm vec}.
\]
The Hilbert-space grading has the opposite sign because the induced
infinitesimal action on local operators contains the differential operator
\(-\mathcal L_\epsilon\), as in \eqref{eq:local-conformal-current-ward}.  In
the radial Hilbert-space representation recalled in
Chapter~\ref{ch:volume-iii-radial-quantization},
\[
  [\widehat D_{\rm rad},\widehat P_\mu]=\widehat P_\mu .
\]
Since \(\widehat D_{\rm rad}\) is self-adjoint and
\(\widehat P_\mu^\dagger=\widehat K_\mu\) by
\eqref{eq:p-k-adjoint-cylinder}, the adjoint commutator is
\([\widehat D_{\rm rad},\widehat K_\mu]=-\widehat K_\mu\).
Consequently \(\widehat P_\mu\) raises and \(\widehat K_\mu\) lowers the
cylinder energy by one unit.

The Hamiltonian identification is also visible directly from the stress
tensor.  On the sphere \(S^{D-1}_r\) of radius \(r\), with \(x^\mu=rn^\mu\),
\begin{equation}
  \widehat D_{\rm rad}
  =
  \int_{S^{D-1}_r}\dd\sigma\,n^\mu \widehat T_{\mu\nu}(x)x^\nu
  =
  r^D\int_{S^{D-1}}\dd\Omega\,
  n^\mu n^\nu \widehat T_{\mu\nu}(rn).
  \label{eq:dilatation-charge-stress-tensor-sphere}
\end{equation}
Under the Weyl map this is the cylinder Hamiltonian
\[
  \widehat H_{\rm cyl}
  =
  \int_{S^{D-1}}\dd\Omega\,\widehat T_{\tau\tau}^{\rm cyl}.
\]
In Lorentzian cylinder time \(t=-\ii\tau\), the corresponding Hermitian
generator differs from the Euclidean radial generator by the standard
Wick-rotation factor described in
Chapter~\ref{ch:volume-iii-radial-quantization}.  The monograph uses
\(\widehat D_{\rm rad}\) for the positive Euclidean cylinder Hamiltonian that
grades local operators by their scaling dimensions; whenever Hermitian
Lorentzian charges appear in Ward identities, the conversion to radial
commutators is made by the same analytic continuation of the group parameter.

\section{Two-Dimensional Chiral Stress-Tensor Charges}
\label{sec:two-dimensional-chiral-stress-tensor-charges}

Two Euclidean dimensions require a refinement of the preceding finite
conformal-charge construction.  Let \(U\) be a coordinate chart on an oriented
Euclidean surface with complex coordinate \(z\).  In a flat chart away from
operator insertions, assume the separated-point stress-tensor Ward identities
take the chiral form
\begin{equation}
  \partial_{\bar z}T(z)=0,\qquad
  \partial_z\bar T(\bar z)=0
  \label{eq:separated-two-dimensional-stress-holomorphy}
\end{equation}
inside correlation functions.  Equation
\eqref{eq:separated-two-dimensional-stress-holomorphy} has the separated
distributional meaning that, for any product \(X\) of local insertions whose
supports avoid \(z\),
\[
  \partial_{\bar z}\langle T(z)X\rangle=0,\qquad
  \partial_z\langle \bar T(\bar z)X\rangle=0 .
\]
Contact terms at the insertions are part of the source convention discussed
in Section~\ref{sec:cft-source-functionals-ward-contacts}.

\begin{definition}[Chiral stress-tensor contour charge]
\label{def:chiral-stress-tensor-contour-charge}
The chiral
contour charge associated with a holomorphic vector field \(v(z)\partial_z\)
on an annulus \(A\) is the oriented contour operator
\begin{equation}
  Q_v(C)
  =
  -\oint_C\frac{\dd z}{2\pi\ii}\,v(z)T(z),
  \label{eq:chiral-contour-charge-vector-field}
\end{equation}
where \(C\subset A\) is positively oriented around the insertion on which the
charge acts.  The sign in \eqref{eq:chiral-contour-charge-vector-field} is
the active-operator sign: the infinitesimal change generated by
\(z\mapsto z+\varepsilon v(z)\) is
\[
  \delta_v\mathcal O
  =
  \varepsilon Q_v(C)\mathcal O .
\]
\end{definition}

Suppose \(\mathcal O(0)\) has holomorphic weight \(h\), so that its response
to translations and linear rescalings is finite and diagonal.  Then the
singular part of the stress-tensor OPE begins as
\begin{equation}
  T(z)\mathcal O(0)
  =
  \cdots
  +
  \frac{h}{z^2}\mathcal O(0)
  +
  \frac1z\partial\mathcal O(0)
  +O(z^0),
  \label{eq:stress-ope-weight-translation-terms}
\end{equation}
where the omitted terms to the left of the displayed terms are possible poles
of order \(z^{-3}\) and higher.  These higher poles are the additional
transformation data under nonlinear chiral vector fields.

\begin{definition}[Virasoro primary and contour modes]
\label{def:virasoro-primary-and-contour-modes}
A Virasoro primary is a local operator \(\phi\) whose stress-tensor singular
part is exactly
\begin{equation}
  T(z)\phi(0)
  =
  \frac{h}{z^2}\phi(0)
  +
  \frac1z\partial\phi(0)
  +O(z^0),
  \qquad
  \bar T(\bar z)\phi(0)
  =
  \frac{\bar h}{\bar z^2}\phi(0)
  +
  \frac1{\bar z}\bar\partial\phi(0)
  +O(\bar z^0).
  \label{eq:virasoro-primary-stress-ope}
\end{equation}
This is stronger than global conformal primarity: it is a statement about the
whole local chiral stress-tensor representation, including the
\(\mathrm{PSL}(2,\mathbb C)\) subalgebra.

On a punctured disc, the holomorphic modes are
\begin{equation}
  L_n
  =
  \oint_0\frac{\dd z}{2\pi\ii}\,z^{n+1}T(z),
  \qquad n\in\mathbb Z,
  \label{eq:virasoro-mode-definition}
\end{equation}
acting first on the common radial domain generated by finitely many local
insertions and then, when reflection positivity and finite-energy
decomposition are available, on the corresponding dense domain in the radial
Hilbert space.  With this convention,
\[
  L_0\ket{\phi}=h\ket{\phi},\qquad
  L_{-1}\ket{\phi}=\ket{\partial\phi}
\]
for a Virasoro primary \(\phi\).  The antiholomorphic modes \(\bar L_n\) are
defined analogously from \(\bar T(\bar z)\).
\end{definition}

\begin{proposition}[Stress-tensor OPE and the Virasoro algebra]
\label{prop:tt-ope-virasoro-algebra}
Assume that the singular part of the holomorphic stress-tensor OPE is
\begin{equation}
  T(z)T(w)
  \sim
  \frac{c}{2(z-w)^4}
  +
  \frac{2T(w)}{(z-w)^2}
  +
  \frac{\partial T(w)}{z-w}.
  \label{eq:stress-tensor-tt-ope}
\end{equation}
Then the contour modes \eqref{eq:virasoro-mode-definition} obey, on the
radial finite-insertion domain,
\begin{equation}
  [L_n,L_m]
  =
  (n-m)L_{n+m}
  +
  \frac{c}{12}(n^3-n)\delta_{n+m,0}.
  \label{eq:virasoro-algebra-from-tt-ope}
\end{equation}
\end{proposition}

\begin{proof}
Let \(C_2\) be a positively oriented contour around the origin and let
\(C_1\) be a positively oriented contour outside \(C_2\).  The ordered product
\(L_nL_m\) is represented by
\[
  \oint_{C_1}\frac{\dd z}{2\pi\ii}
  \oint_{C_2}\frac{\dd w}{2\pi\ii}\,
  z^{n+1}w^{m+1}T(z)T(w).
\]
The product \(L_mL_n\) is the same expression with the \(z\)-contour moved to
a contour \(C'_1\) inside \(C_2\).  Hence the commutator is obtained by
integrating \(z\) over \(C_1-C'_1\).  Since \(C_1-C'_1\) encloses the point
\(w\) but not the origin, it can be shrunk to a small positive contour around
\(w\).  Only the singular part of \eqref{eq:stress-tensor-tt-ope}
contributes:
\[
  [L_n,L_m]
  =
  \oint_{C_2}\frac{\dd w}{2\pi\ii}
  w^{m+1}
  \operatorname*{Res}_{z=w}
  \left[
    z^{n+1}
    \left(
      \frac{c}{2(z-w)^4}
      +
      \frac{2T(w)}{(z-w)^2}
      +
      \frac{\partial T(w)}{z-w}
    \right)
  \right].
\]
The residues are
\[
  \operatorname*{Res}_{z=w}\frac{z^{n+1}}{(z-w)^4}
  =
  \frac{(n+1)n(n-1)}6\,w^{n-2},
  \qquad
  \operatorname*{Res}_{z=w}\frac{z^{n+1}}{(z-w)^2}
  =
  (n+1)w^n .
\]
Thus
\[
  [L_n,L_m]
  =
  \oint_{C_2}\frac{\dd w}{2\pi\ii}
  \left[
    \frac{c}{12}(n^3-n)w^{m+n-1}
    +
    2(n+1)w^{m+n+1}T(w)
    +
    w^{m+n+2}\partial T(w)
  \right].
\]
The first term contributes precisely when \(m+n=0\).  For the last term,
integrating by parts on the closed contour gives
\[
  \oint\frac{\dd w}{2\pi\ii}\,
  w^{m+n+2}\partial T(w)
  =
  -(m+n+2)
  \oint\frac{\dd w}{2\pi\ii}\,
  w^{m+n+1}T(w).
\]
The coefficient of the remaining stress-tensor mode is therefore
\(2(n+1)-(m+n+2)=n-m\), giving
\eqref{eq:virasoro-algebra-from-tt-ope}.
\end{proof}

\begin{remark}[Antiholomorphic sector and gravitational anomaly]
The antiholomorphic modes satisfy the same algebra with \(c\) replaced by
\(\bar c\), and separated holomorphic and antiholomorphic contours commute up
to possible contact terms that would be a gravitational anomaly.  A
reflection-positive parity-invariant two-dimensional CFT has \(c=\bar c\).
In a theory with a gravitational anomaly, the two chiral central charges are
different and the spatial momentum on the cylinder acquires the corresponding
vacuum shift.
\end{remark}

\paragraph{Infinitesimal stress-tensor transformation.}
With the contour-charge convention
\eqref{eq:chiral-contour-charge-vector-field} and the stress-tensor OPE
\eqref{eq:stress-tensor-tt-ope}, a holomorphic vector field
\(v(z)\partial_z\) acts on \(T(w)\) by
\begin{equation}
  \delta_v T(w)
  =
  -v(w)\partial T(w)-2v'(w)T(w)-\frac{c}{12}v'''(w).
  \label{eq:infinitesimal-stress-tensor-transform}
\end{equation}
This is an equality of insertions on any annulus on which the contour for
\(Q_v\) can be chosen without crossing other operators.  The formula is
therefore a separated-point Ward identity; contact extensions at collisions
are fixed by the source convention for the stress tensor.

The contour calculation is local around \(w\).  By definition,
\[
  \delta_vT(w)=Q_v(C)T(w)
  =
  -\oint_C\frac{\dd z}{2\pi\ii}\,v(z)T(z)T(w),
\]
where \(C\) is a small positive contour around \(w\).  Insert
\eqref{eq:stress-tensor-tt-ope}.  The three required residues are
\[
  \operatorname*{Res}_{z=w}\frac{v(z)}{z-w}=v(w),
  \qquad
  \operatorname*{Res}_{z=w}\frac{v(z)}{(z-w)^2}=v'(w),
  \qquad
  \operatorname*{Res}_{z=w}\frac{v(z)}{(z-w)^4}=\frac{v'''(w)}6 .
\]
The overall minus sign in \(Q_v\) gives
\[
  -\left(
    v(w)\partial T(w)
    +2v'(w)T(w)
    +\frac c2\,\frac{v'''(w)}6
  \right),
\]
which is \eqref{eq:infinitesimal-stress-tensor-transform}.

\paragraph{Schwarzian cocycle.}
For locally univalent holomorphic maps \(f\) and \(g\), the Schwarzian
derivative
\[
  \{f,z\}
  =
  \frac{f'''(z)}{f'(z)}
  -
  \frac32\left(\frac{f''(z)}{f'(z)}\right)^2
\]
satisfies
\begin{equation}
  \{g\circ f,z\}
  =
  (f'(z))^2\{g,f(z)\}+\{f,z\}.
  \label{eq:schwarzian-cocycle}
\end{equation}
Moreover, if \(f_s(z)=z+s v(z)+O(s^2)\), then
\begin{equation}
  \left.\frac{\dd}{\dd s}\right|_{s=0}\{f_s,z\}
  =
  v'''(z).
  \label{eq:schwarzian-linearization}
\end{equation}

Set
\[
  A_f(z)=\frac{f''(z)}{f'(z)}=\partial_z\log f'(z).
\]
Then
\[
  \{f,z\}=A_f'(z)-\frac12 A_f(z)^2.
\]
For a composition,
\[
  A_{g\circ f}(z)
  =
  A_f(z)+f'(z)A_g(f(z)).
\]
Differentiate and subtract one half of the square:
\[
\begin{aligned}
  A_{g\circ f}'-\frac12A_{g\circ f}^2
  &=
  A_f'-\frac12A_f^2
  +(f')^2\left(A_g'-\frac12A_g^2\right)\!\circ f
\\
  &\qquad
  +f''A_g\circ f-A_f f' A_g\circ f .
\end{aligned}
\]
Since \(A_f=f''/f'\), the last two terms cancel.  This proves
\eqref{eq:schwarzian-cocycle}.  Finally, for
\(f_s(z)=z+s v(z)+O(s^2)\),
\[
  f_s'=1+s v'+O(s^2),\qquad
  f_s''=s v''+O(s^2),\qquad
  f_s'''=s v'''+O(s^2).
\]
Substitution into the definition of the Schwarzian gives
\(\{f_s,z\}=s v'''(z)+O(s^2)\), proving
\eqref{eq:schwarzian-linearization}.

\begin{proposition}[Schwarzian transformation and plane-cylinder shift]
\label{prop:schwarzian-transformation-plane-cylinder-shift}
The stress tensor itself carries an affine chiral transformation law fixed by
\eqref{eq:stress-tensor-tt-ope}.  If \(z'=f(z)\) is a holomorphic coordinate
change, then
\begin{equation}
  T'(z')
  =
  \left(\frac{\dd f}{\dd z}\right)^{-2}
  \left[
    T(z)-\frac{c}{12}\{f,z\}
  \right],
  \qquad
  \{f,z\}
  =
  \frac{f'''(z)}{f'(z)}
  -
  \frac32
  \left(\frac{f''(z)}{f'(z)}\right)^2 .
  \label{eq:stress-tensor-schwarzian-transformation}
\end{equation}
The Schwarzian derivative satisfies the cocycle identity
\[
  \{g\circ f,z\}
  =
  (f'(z))^2\{g,f(z)\}+\{f,z\},
\]
which is exactly the condition that successive coordinate changes act
associatively on \(T\).

For the cylinder coordinate \(w=\sigma+\ii\tau\), \(\sigma\sim\sigma+2\pi\),
we use
\[
  z=\ee^{-\ii w},
  \qquad
  w=\ii\log z.
\]
Applying \eqref{eq:stress-tensor-schwarzian-transformation} to the coordinate
change \(z\mapsto w(z)\), for which \(\{w,z\}=1/(2z^2)\), gives
\begin{equation}
  T^{(w)}(w)
  =
  \left(\frac{\dd z}{\dd w}\right)^2T^{(z)}(z)
  +
  \frac{c}{24}
  =
  -z^2T^{(z)}(z)+\frac{c}{24}.
  \label{eq:plane-cylinder-stress-tensor-shift}
\end{equation}
If
\[
  T^{(z)}(z)=\sum_{n\in\mathbb Z}\frac{L_n}{z^{n+2}},
  \qquad
  \bar T^{(\bar z)}(\bar z)=
  \sum_{n\in\mathbb Z}\frac{\bar L_n}{\bar z^{n+2}},
\]
then
\[
  T^{(w)}(w)
  =
  -\sum_{n\in\mathbb Z}\ee^{\ii n w}L_n+\frac{c}{24},
  \qquad
  \bar T^{(\bar w)}(\bar w)
  =
  -\sum_{n\in\mathbb Z}\ee^{-\ii n\bar w}\bar L_n
  +\frac{\bar c}{24}.
\]
In the cylinder normalization used in
Chapter~\ref{ch:volume-iii-radial-quantization}, the Hamiltonian and spatial
momentum on the unit circle are therefore
\begin{equation}
  H_{\rm cyl}
  =
  L_0+\bar L_0-\frac{c+\bar c}{24},
  \qquad
  P_{\rm cyl}
  =
  L_0-\bar L_0-\frac{c-\bar c}{24}.
  \label{eq:cylinder-hamiltonian-momentum-virasoro}
\end{equation}
Thus the radial Hamiltonian differs from the naive dilatation eigenvalue by
the Casimir term forced by the stress-tensor anomaly under the plane-cylinder
coordinate change.
\end{proposition}

\begin{proof}
First prove the finite formula.  For a local coordinate change \(f\), define
the right side of \eqref{eq:stress-tensor-schwarzian-transformation} by
\[
  \mathcal T_f(f(z))
  =
  (f'(z))^{-2}
  \left[T(z)-\frac c{12}\{f,z\}\right].
\]
If \(f_s(z)=z+s v(z)+O(s^2)\), then
\[
  (f_s')^{-2}=1-2s v'+O(s^2),
  \qquad
  T(z)=T(z_s-sv(z_s))+O(s^2)
\]
at the fixed final point \(z_s=f_s(z)\).  Together with
\eqref{eq:schwarzian-linearization},
\[
  \mathcal T_{f_s}(z_s)
  =
  T(z_s)
  -s\left(
    v\partial T+2v'T+\frac c{12}v'''
  \right)(z_s)
  +O(s^2),
\]
which is exactly the infinitesimal Ward action
\eqref{eq:infinitesimal-stress-tensor-transform}.

It remains to check compatibility with finite composition.  Applying first
\(f\) and then \(g\) gives
\[
  (g'(f)f')^{-2}
  \left[
    T
    -\frac c{12}\{f,z\}
    -\frac c{12}(f')^2\{g,f(z)\}
  \right],
\]
which equals the transformation for \(g\circ f\) precisely by
\eqref{eq:schwarzian-cocycle}.  Thus the displayed formula is a
projective cocycle for the connected local conformal group whose tangent
action is the stress-tensor Ward action.  On a simply connected
neighborhood of the identity this characterizes the integrated action
uniquely; analytic continuation along paths in the chosen coordinate-change
group gives the stated finite law, with possible global monodromy recorded
separately by the gravitational-anomaly data.

For \(w=\ii\log z\), direct differentiation gives
\[
  w'=\frac{\ii}{z},\qquad
  w''=-\frac{\ii}{z^2},\qquad
  w'''=\frac{2\ii}{z^3},
  \qquad
  \{w,z\}=\frac{1}{2z^2}.
\]
Using \(z=\ee^{-\ii w}\) and \(\dd z/\dd w=-\ii z\) in
\eqref{eq:stress-tensor-schwarzian-transformation} yields
\eqref{eq:plane-cylinder-stress-tensor-shift}.  Substituting the Laurent
expansions of \(T^{(z)}\) and \(\bar T^{(\bar z)}\) and taking the zero modes
of cylinder translations and rotations gives
\eqref{eq:cylinder-hamiltonian-momentum-virasoro}.
\end{proof}

\begin{remark}[Local operators and the radial Hilbert completion in \(D=2\)]
The Virasoro mode construction refines the domain statement of
Chapter~\ref{ch:volume-iii-radial-quantization}.  The local
operators at the origin, including their Virasoro descendants, span an
algebraic finite-energy subspace.  After quotienting null vectors and
completing in the BPZ/radial norm, this subspace is dense in the Hilbert
space on the circle under the hypotheses of
Hypothesis~\ref{hyp:radial-reconstruction-data}.  The precise
state--operator statement is therefore the correspondence between local
scaling fields of definite \(L_0,\bar L_0\) eigenvalues and the finite-energy
radial eigenstates in this algebraic domain.  The completed Hilbert space is
the BPZ/radial completion of that dense domain.
This density is the analytic input behind the radial OPE: the product of two
nearby local insertions is represented by a convergent expansion in local
scaling fields in the radial Hilbert norm.
\end{remark}
