\chapter{The Operator Product Expansion}
\label{ch:volume-iii-operator-product-expansion}
\ConventionsEuclideanCFT


\section{Local Expansion in Radial Quantization}

Let \(\mathcal O_i(x)\) and \(\mathcal O_j(0)\) be local operators inserted
inside a sphere of radius \(R\), and let all other insertions lie outside that
sphere.  Radial quantization turns the separated pair
\[
  \Psi_{ij}(x):=\mathcal O_i(x)\mathcal O_j(0)\ket{\vac},
  \qquad |x|<R,
\]
into a state on \(S_R^{D-1}\).  This state is not, as an object, a local
operator at the origin, and there is no operation in the local-operator
space which sends the ordered pair
\((\mathcal O_i(x),\mathcal O_j(0))\), with \(x\ne0\), to a new local
operator at \(0\).  The local-operator states of definite scaling dimension
form an algebraic dense subspace of the BPZ/radial Hilbert space, and the
finite-energy spectral projections of the separated-pair state are finite
linear combinations of those local-operator states.  If \(P_{\le E}\) denotes
the spectral projection of \(D_{\rm rad}\) onto scaling dimensions at most
\(E\), the OPE is the Hilbert-norm convergence
\[
  \lim_{E\to\infty}P_{\le E}\Psi_{ij}(x)=\Psi_{ij}(x).
\]
Written in local-operator coordinates on the dense finite-energy subspace,
the same statement is conventionally abbreviated as
\begin{equation}
  \Psi_{ij}(x)
  \sim_{\rm BPZ}
  \sum_{\mathcal O}
  C_{ij\mathcal O}(x,\partial)\,\mathcal O(0).
  \label{eq:ope-general-local}
\end{equation}
The sum runs over primary operators \(\mathcal O\) and the differential
operator \(C_{ij\mathcal O}(x,\partial)\) generates the descendants of
\(\mathcal O\).  The symbol \(\sim_{\rm BPZ}\) does not denote equality in a
pointwise algebra of local operators.  It denotes that the finite partial
sums over dimensions \(\Delta\le\Delta_*\) define local-operator states at the
origin whose sequence converges to \(\Psi_{ij}(x)\) in the BPZ/radial Hilbert
norm on the separating sphere, and hence inside correlation functions with
all other insertions farther from the origin than \(|x|\).

Equivalently, write \(x=\ee^\tau n_x\) under the Weyl map
\(\R^D\setminus\{0\}\simeq\R_\tau\times S^{D-1}\).  For scalar primaries,
\(\mathcal O_i(x)=|x|^{-\Delta_i}\widetilde{\mathcal O}_i(\tau,n_x)\),
so the product creates the cylinder state
\[
  \widetilde{\mathcal O}_i(\tau,n_x)\ket{\mathcal O_j}
  \in\mathcal H_{S^{D-1}}.
\]
The OPE is the statement that this state is represented, in every exterior
correlation function, by a convergent limit of local operators at the origin,
organized into primary states and their descendants.  Thus the expansion is
not a formal Taylor series in elementary fields and not a multiplication law
closing the local-operator space.  It is the spectral expansion of the
Hilbert-space state created by separated local insertions by a sequence in
the dense algebraic subspace of local-operator states.

The convergence parameter is radial separation.  If the nearest outside
insertion is at radial distance \(R\), the expansion is organized by powers of
\(|x|/R\), with larger scaling dimensions contributing at higher radial
energy.  The OPE is therefore a convergent spectral expansion controlled by
the local operator spectrum and the conformal representation theory of
descendants.

\begin{definition}[Radial OPE inputs from radial reconstruction]
\label{def:radial-ope-convergence-hypotheses}
The Euclidean radial OPE theorem below is stated for a Schwinger hierarchy
satisfying the radial reconstruction data of
Hypothesis~\ref{hyp:radial-reconstruction-data}.  The following list is not an
independent set of assumptions; it records which clauses of that single
radial-reconstruction hypothesis enter the Euclidean convergence proof:
\begin{enumerate}
  \item distributional Euclidean correlators of renormalized local operators,
  with covariance, graded permutation symmetry, and contact-term
  prescriptions;
  \item radial reflection positivity and the corresponding null quotient and
  Hilbert completion;
  \item a nonnegative self-adjoint radial Hamiltonian \(D_{\rm rad}\) with
  discrete point spectrum and finite-dimensional spectral subspaces below
  every finite energy;
  \item radial local completeness, so the algebraic direct sum of local
  scaling-operator states is dense and its homogeneous components identify
  with the \(D_{\rm rad}\)-eigenspaces as in
  Theorem~\ref{thm:state-operator-correspondence}.
\end{enumerate}
The Lorentzian continuation statements in this chapter additionally use the
tube-domain analytic continuation, spectrum condition, and local
commutativity part of the same hypothesis.  Thus the phrase
``radial OPE convergence under reflection positivity'' denotes this listed
Hilbert-space and spectral package.
\end{definition}

\begin{theorem}[Euclidean radial OPE convergence]
\label{thm:euclidean-radial-ope-convergence}
  Assume Hypothesis~\ref{hyp:radial-reconstruction-data}.  Let
  \(\mathcal O_i(x_E)\mathcal O_j(0)\vac\) be inserted at Euclidean points
  inside a sphere of radius \(r\) in \(\R^D_E\), and let all remaining
  Euclidean insertions lie outside radius \(R>r\).  Then the expansion of this
  state in eigenspaces of the dilatation operator converges in Hilbert norm for
  \(r<R\), and the resulting OPE converges inside the corresponding Euclidean
  correlation functions.  Truncating to states of scaling dimension at most
  \(\Delta_*\) gives a remainder controlled by powers of \(r/R\) determined by
  the omitted dimensions.
\end{theorem}

The proof is the spectral theorem applied to Euclidean radial quantization.
Radial time is \(\tau=\log |x_E|\), and the dilatation operator is the
nonnegative self-adjoint radial Hamiltonian \(D_{\rm rad}\) of
Hypothesis~\ref{hyp:radial-reconstruction-data}.  A separating sphere at radius
\(R\) inserts the damping factor
\[
  \exp[-(\log(R/r))D_{\rm rad}] .
\]
Since the spectrum is discrete with finite multiplicities below any fixed
dimension, the spectral projection onto \(\Delta\le\Delta_*\) gives a finite
partial OPE and the complement is suppressed by the corresponding power of
\(r/R\) relative to the norm of the external state.  This is the mathematical
reason the Euclidean OPE is an operator algebraic expansion in unitary CFTs
satisfying the stated radial hypotheses.

The preceding theorem is a statement in the Euclidean radial Hilbert space.
It does not by itself assert pointwise convergence at arbitrary real
Lorentzian configurations.  Lorentzian OPE convergence is obtained by applying
the Euclidean expansion inside a common holomorphic domain and then taking the
appropriate Wightman boundary value.

\begin{proposition}[Lorentzian continuation of the radial OPE]
\label{prop:lorentzian-ope-continuation-domain}
Assume the tube-domain part of
Hypothesis~\ref{hyp:radial-reconstruction-data}.  Work with a fixed Wightman
ordering \(\pi\) and write \(V_+\subset\R^{1,D-1}\) for the open future light
cone in the mostly-plus metric.  Let
\(\mathcal W_\pi(z_1,\ldots,z_n)\) be the holomorphic continuation in the tube
component
\[
  z_{\pi(a)}-z_{\pi(a+1)}
  =
  \xi_a-\ii\eta_a,
  \qquad
  \eta_a\in V_+ .
\]
For a chosen OPE channel, let \(\rho,\bar\rho\) denote the radial cross-ratios
of that channel.  On every compact subset of the tube component on which
\[
  |\rho|\le q,\qquad |\bar\rho|\le q,\qquad q<1,
  \label{eq:lorentzian-rho-compact-domain}
\]
the channel OPE converges locally uniformly as a holomorphic function.  Its
boundary value is the Lorentzian Wightman OPE in that ordering.  Strictly
timelike configurations in which the fused pair lies in each other's future or
past light cone are included when the specified \(i\epsilon\) approach remains
in such a compact radial domain.  Null separation, coincident points, and
crossings of external light cones are boundary strata and require the
distributional boundary-value interpretation.
\end{proposition}

\begin{proof}
The Euclidean radial theorem gives a convergent expansion in the open polydisc
of radial variables.  On a compact set satisfying
\eqref{eq:lorentzian-rho-compact-domain}, the same spectral estimate supplies a
summable majorant: each level-\(\Delta\) descendant contribution is multiplied
by a factor bounded by \(q^\Delta\), while finite multiplicity below each level
and radial Hilbert convergence control the corresponding coefficient norm.
Therefore the OPE partial sums converge uniformly on compact subsets of the
radial polydisc after analytic continuation of \(\rho,\bar\rho\).

Each finite partial sum is a finite linear combination of Wightman matrix
elements of local operators at the origin with the remaining insertions.  By
the spectrum condition, these matrix elements are boundary values of functions
holomorphic in the tube component displayed above.  Local uniform convergence
of holomorphic partial sums gives a holomorphic limit by the Weierstrass
theorem.  On the Euclidean subdomain this limit equals
\(\mathcal W_\pi\), because both sides are the Euclidean radial OPE.  The
identity theorem then identifies the holomorphic limit with
\(\mathcal W_\pi\) throughout the connected tube component reached without
crossing singular hypersurfaces.  Taking ordered \(i\epsilon\) boundary values
gives the stated Lorentzian distributional convergence.
\end{proof}

\begin{remark}[Status of Lorentzian OPE convergence]
The criterion above separates the Euclidean Hilbert-space theorem from its
Lorentzian continuation.  The tube condition fixes the operator ordering, and
the radial inequality \(|\rho|,|\bar\rho|<1\) fixes the OPE channel.  For
scalar four-point functions in unitary Euclidean CFT, Kravchuk, Qiao, and
Rychkov prove the needed forward-tube radial bound and distributional
Minkowski OPE convergence; Qiao gives a channel-by-channel classification of
real Lorentzian four-point configurations.\footnote{See G.~Mack,
``Convergence of operator product expansions on the vacuum in conformal
invariant quantum field theory,'' Commun. Math. Phys. \textbf{53} (1977)
155--184; P.~Kravchuk, J.~Qiao, and S.~Rychkov, ``Distributions in CFT II.
Minkowski Space,'' JHEP \textbf{08} (2021) 094, arXiv:2104.02090; and
J.~Qiao, ``Classification of Convergent OPE Channels for Lorentzian CFT
Four-Point Functions,'' SciPost Phys. \textbf{13} (2022) 093,
arXiv:2005.09105.}
\end{remark}

\begin{figure}[H]
  \centering
  \resizebox{0.94\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      insertion/.style={circle,fill=violet!80!black,inner sep=1.7pt},
      arrow/.style={-{Latex[length=2mm]},line width=0.8pt},
      surface/.style={green!45!black,line width=0.8pt}
    }
    \begin{scope}
      \node[font=\small] at (0,2.0) {two nearby insertions};
      \draw[thick] (0,0) circle (1.45);
      \draw[surface] (0,0) circle (0.92);
      \node[insertion] at (-0.28,0.10) {};
      \node[insertion] at (0.35,0.35) {};
      \node[violet!80!black] at (-0.58,-0.20) {\(\mathcal O_j(0)\)};
      \node[violet!80!black] at (0.82,0.55) {\(\mathcal O_i(x)\)};
      \node[green!45!black] at (1.12,-0.70) {\(S_R^{D-1}\)};
      \node[blue!65!black] at (0,-1.92) {outside insertions are outside \(S_R^{D-1}\)};
    \end{scope}

    \draw[arrow] (2.00,0.0)--(3.25,0.0);
    \node[align=center] at (2.62,0.52) {expand\\on sphere};

    \begin{scope}[shift={(5.55,0)}]
      \node[font=\small] at (0,2.0) {primary families};
      \draw[thick] (0,0) circle (1.45);
      \draw[surface] (0,0) circle (0.92);
      \foreach \y/\lab in {0.66/{\mathcal O},0.22/{\partial\mathcal O},-0.22/{\partial^2\mathcal O},-0.66/{\cdots}} {
        \node[insertion] at (0,\y) {};
        \node[violet!80!black,right] at (0.18,\y) {\(\lab\)};
      }
      \node[orange!85!black] at (0,-1.85)
        {\(\displaystyle \sum_{\mathcal O} C_{ij\mathcal O}(x,\partial)\mathcal O(0)\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The OPE is the expansion of the state created by nearby insertions
  in the basis of primary conformal multiplets on a separating sphere.}
  \label{fig:ope-radial-expansion}
\end{figure}

\section{Scalar-Scalar OPE}

Let \(\phi_1\) and \(\phi_2\) be scalar primaries of dimensions
\(\Delta_1\) and \(\Delta_2\).  Let
\(\mathcal O_{\mu_1\cdots\mu_\ell}\) be a symmetric traceless primary of spin
\(\ell\) and dimension \(\Delta\).  Its leading contribution to the
\(\phi_1\times\phi_2\) OPE has the following state-valued asymptotic
coordinate in the BPZ/radial expansion:
\begin{equation}
  \phi_1(x)\phi_2(0)
  \supset
  \lambda_{12\mathcal O}
  \frac{x^{\mu_1}\cdots x^{\mu_\ell}}
       {|x|^{\Delta_1+\Delta_2-\Delta+\ell}}
  \mathcal O_{\mu_1\cdots\mu_\ell}(0)
  +\text{descendants}.
  \label{eq:scalar-scalar-leading-ope}
\end{equation}
The tensor \(x^{\mu_1}\cdots x^{\mu_\ell}\) selects the symmetric traceless
part of the spin-\(\ell\) representation.  Antisymmetric components do not
appear in the product of powers of a single vector, and trace parts are
absorbed into descendants or lower-spin structures.

For an exchanged scalar primary \(\mathcal O\), the first descendant term is
fixed by conformal symmetry:
\begin{equation}
  \phi_1(x)\phi_2(0)
  \supset
  \lambda_{12\mathcal O}
  |x|^{\Delta-\Delta_1-\Delta_2}
  \left(
    1+
    \frac{\Delta+\Delta_1-\Delta_2}{2\Delta}
    x^\mu\partial_\mu+\cdots
  \right)\mathcal O(0).
  \label{eq:scalar-ope-first-descendant}
\end{equation}
The coefficient of the derivative is determined by requiring the OPE to
reproduce the conformally fixed three-point function
\(\langle\phi_1\phi_2\mathcal O\rangle\).  To see the coefficient, place the
third operator at \(z\) with \(|x|\ll |z|\) and normalize
\(\langle\mathcal O(y)\mathcal O(z)\rangle=|y-z|^{-2\Delta}\).  The
three-point function gives
\[
  \langle\phi_1(x)\phi_2(0)\mathcal O(z)\rangle
  =
  \lambda_{12\mathcal O}
  |x|^{\Delta-\Delta_1-\Delta_2}|z|^{-2\Delta}
  \left[
    1+
    (\Delta+\Delta_1-\Delta_2)
    \frac{x\cdot z}{z^2}
    +O(x^2)
  \right].
\]
The OPE ansatz
\[
  \lambda_{12\mathcal O}
  |x|^{\Delta-\Delta_1-\Delta_2}
  \bigl(1+a\,x^\mu\partial_\mu+\cdots\bigr)\mathcal O(0)
\]
gives instead
\[
  \lambda_{12\mathcal O}
  |x|^{\Delta-\Delta_1-\Delta_2}|z|^{-2\Delta}
  \left[
    1+2a\Delta\frac{x\cdot z}{z^2}+O(x^2)
  \right],
\]
because
\(\partial_{y^\mu}|y-z|^{-2\Delta}|_{y=0}
=2\Delta z_\mu |z|^{-2\Delta-2}\).  Matching the two expansions gives
\[
  a=\frac{\Delta+\Delta_1-\Delta_2}{2\Delta}.
\]
Higher descendants are fixed in the same way by matching higher orders, or
equivalently by the conformal algebra.

For identical scalar primaries this reduces to the especially important
normalization
\begin{equation}
  \phi(x)\phi(0)
  \supset
  \lambda_{\phi\phi\mathcal O}
  |x|^{\Delta_{\mathcal O}-2\Delta_\phi}
  \left(1+\frac12 x^\mu\partial_\mu+\cdots\right)\mathcal O(0)
  \label{eq:identical-scalar-scalar-ope-first-descendant}
\end{equation}
for scalar exchange.  The coefficient \(1/2\) is not an extra convention; it
is the specialization of the conformal Ward constraints to two equal external
dimensions.

\section{OPE Data}

Choose a basis of primaries with diagonal two-point functions.  The local CFT
data are then
\begin{equation}
  \left\{
    \Delta_{\mathcal O},\;
    \rho_{\mathcal O},\;
    \lambda_{ij\mathcal O}^{(a)}
  \right\}.
  \label{eq:cft-local-ope-data}
\end{equation}
Here \(\rho_{\mathcal O}\) is the spin representation, and the label \(a\)
enumerates independent three-point tensor structures when more than one
structure exists.  Once this data is specified, all descendant coefficients in
the OPE are representation-theoretic consequences of the conformal algebra.

In representation language, if \(V_i\) is the conformal multiplet generated
by a primary \(\mathcal O_i\), the OPE map is a conformally covariant
state-valued intertwining map from the tensor product of two local-state
multiplets, together with a nonzero radial separation, to the Hilbert space
on the separating sphere.  On the decomposition into irreducible conformal
multiplets its finite-energy projections have the form
\[
  V_i\otimes V_j
  \longrightarrow
  \bigoplus_{\mathcal O} m_{ij}^{\mathcal O}V_{\mathcal O}
  \subset \mathcal H_{S^{D-1}} .
\]
The multiplicity \(m_{ij}^{\mathcal O}\) equals the number of independent
three-point structures connecting \(i\), \(j\), and \(\mathcal O\).  When
\(\mathcal O_i\) and \(\mathcal O_j\) are both scalars, rotational covariance
permits only symmetric traceless tensor primaries in the leading term, and
the multiplicity for each such representation is one.  For spinning external
operators, several independent tensor structures may couple the same three
representations, and the OPE coefficients must then carry the corresponding
structure label.

Repeated OPEs determine every separated Euclidean correlation function from
the primary spectrum, the two-point pairing, and the three-point
coefficients, provided the radial convergence assumptions stated above hold.
The data are therefore local in a strong sense: higher-point functions do not
bring in independent local constants.  Their nontrivial content is the
compatibility of the same OPE data in different channels.

\section{Reconstruction from OPE Data}
\label{sec:cft-reconstruction-from-ope-data}

The preceding paragraph has a converse direction only after the word
``data'' is given a precise mathematical meaning.  A complete primary OPE
table
\(\{\Delta_{\mathcal O},\rho_{\mathcal O},\lambda_{ij\mathcal O}^{(a)}\}\),
with all primary labels, degeneracies, tensor structures, and operator-basis
conventions resolved, is the table of radial expansion coefficients for
separated-pair states in the dense local-operator subspace.  It is therefore
a candidate for an inverse determination of the theory, in the same sense
that a sufficiently rich exact set of low-energy observables can in principle
determine the parameters and operator content of an effective theory.  If a
smaller set of local operators generates the finite-energy local-operator
subspace under iterated radial OPE limits, the complete OPE spectrum and
coefficients visible in suitable four-point functions of those generators may
likewise determine the rest of the local theory.  This rigidity is not, by
itself, a reconstruction theorem.  A reconstruction theorem must also specify
the Hilbert topology, positivity, radial convergence, channel compatibility,
and the distributional
interpretation of separated correlators.  The following formulation is the
radial CFT counterpart of a Wightman or Osterwalder--Schrader reconstruction
theorem.

\begin{definition}[Abstract radial OPE datum]
\label{def:abstract-radial-ope-datum}
An abstract radial OPE datum in dimension \(D\ge2\) consists of the following
objects.
\begin{enumerate}[label=(\roman*)]
  \item A countable set \(\mathcal P\) of primary labels, including the
  identity label \(\mathbf 1\).  Each \(a\in\mathcal P\) carries a scaling
  dimension \(\Delta_a\in\R_{\ge0}\) and a finite-dimensional unitary
  representation \(\rho_a\) of \(\operatorname{Spin}(D)\).  For every
  \(E<\infty\), only finitely many descendant states in the modules below have
  dimension at most \(E\).
  \item For each \(a\), a positive-energy conformal module
  \(V_a\) generated by a primary vector \(\ket a\), with null descendants
  quotiented.  The direct sum
  \[
    \mathcal H_{\rm fin}^{\rm abs}
    =
    \bigoplus_{a\in\mathcal P} V_a
  \]
  is equipped with a positive Hermitian inner product for which the conformal
  generators obey the radial adjointness relations
  \(D^\dagger=D\), \(M_{\mu\nu}^\dagger=-M_{\mu\nu}\), and
  \(P_\mu^\dagger=K_\mu\).
  \item A conjugation \(a\mapsto a^\dagger\) and a nondegenerate two-point
  pairing between \(V_a\) and \(V_{a^\dagger}\), normalized so that
  \(V_{\mathbf 1}\) is the one-dimensional vacuum module.
  \item For every ordered pair \(a,b\) and every \(c\in\mathcal P\), a finite
  list of conformally covariant three-point tensor structures
  \(\mathcal T_{ab}^{c,\alpha}\), \(\alpha=1,\ldots,m_{ab}^{c}\), and
  complex coefficients \(\lambda_{ab}^{c,\alpha}\).  The descendant OPE
  coefficients are the coefficients obtained by applying the conformal algebra
  and the quotient Gram matrices inside \(V_c\).
  \item For every finite ordered list of primary or descendant labels and every
  rooted binary OPE tree \(T\), an iterated OPE series
  \[
    S_T(x_1,\ldots,x_n)
  \]
  on the radial configuration domain of \(T\).  This series is required to
  converge in the radial Hilbert norm for state-valued partial products and
  locally uniformly, after pairing with exterior finite-energy states, on
  compact subsets of the separated-point domain of \(T\).
  \item Channel compatibility: if two trees \(T\) and \(T'\) have overlapping
  radial domains, the functions \(S_T\) and \(S_{T'}\) agree on the common
  domain after the canonical permutation and tensor-index identifications.
  For four-point functions this condition is crossing, namely equality of two
  binary radial-expansion limits tested between exterior states.  For
  higher-point functions it is the equality of all iterated OPE
  parenthesizations on their common radial domains.  A separate coherence
  theorem may allow the higher-point identities to be derived from a smaller
  list of associativity identities; absent such a theorem, the all-tree
  compatibility is part of the datum.
  \item Euclidean covariance under the conformal group, graded permutation
  symmetry at separated points, and radial reflection positivity for every
  finite linear combination of inside-supported products.
  \item A prescription for extending the separated-point functions to
  distributions with contact terms on diagonals, whenever coincident-point
  Ward identities or deformations are part of the theory.
\end{enumerate}
\end{definition}

The finite-multiplicity condition in item (i) is part of the datum because
the radial OPE theorem uses spectral projections onto finite-dimensional
energy windows.  The compatibility condition in item (vi) is also part of the
datum.  A single four-point crossing equation is one instance of this
condition and is normally far from determining even that four-point function
uniquely.  Complete all-primary four-point associativity constraints are a
different object: once the OPE table is known, they constrain the whole
system of binary radial-expansion limits.  The theorem below nevertheless
assumes the analytic and higher-coherence properties needed for
reconstruction, rather than renaming their desired conclusion as an input
theorem.

\begin{theorem}[Conditional radial reconstruction from OPE data]
\label{thm:conditional-cft-reconstruction-from-ope}
Let \(\mathfrak D\) be an abstract radial OPE datum in the sense of
Definition~\ref{def:abstract-radial-ope-datum}.  Then \(\mathfrak D\)
determines a separated-point Euclidean conformal correlation hierarchy
\[
  S_{I_1\cdots I_n}^{\mathfrak D}(x_1,\ldots,x_n),
  \qquad x_i\ne x_j,
\]
unique up to unitary equivalence of the radial Hilbert space and invertible
changes of local operator basis preserving the two-point pairing.  This
hierarchy satisfies Euclidean conformal covariance, graded permutation
symmetry, radial reflection positivity, the spectral discreteness and
finite-multiplicity conditions of radial quantization, and the OPE with the
prescribed primary dimensions, spin representations, and three-point
coefficients.  If the datum also includes tube-domain analytic continuation,
the spectrum condition, and local commutativity for the boundary values, then
the reconstructed Euclidean hierarchy gives a Lorentzian Wightman CFT
satisfying the radial reconstruction hypotheses used in this volume.
\end{theorem}

\begin{proof}
First construct the dense finite-energy vector space
\(\mathcal H_{\rm fin}^{\rm abs}=\bigoplus_a V_a\) with the positive inner
product specified in Definition~\ref{def:abstract-radial-ope-datum}.  Its
Hilbert completion is the radial Hilbert space.  The conformal generators act
on each \(V_a\), and the adjointness relations make the representation
unitary on the finite-energy domain.  The identity module supplies the vacuum
vector.

For a separated Euclidean configuration, choose a rooted binary OPE tree whose
radial inequalities are satisfied by the insertion points.  Apply the
intertwining OPE maps at the internal vertices of the tree and sum over
intermediate primary modules and descendants.  Item (v) gives a convergent
state-valued expansion at each internal radial step and locally uniform
convergence after pairing with the remaining exterior states.  Hence the tree
defines a correlation function \(S_T(x_1,\ldots,x_n)\) on its radial domain.

If two admissible trees describe the same configuration on an overlap, item
(vi) identifies the two sums there.  The separated Euclidean configuration
space is covered by such radial domains.  Patching the locally defined
functions through their overlaps gives a single separated-point hierarchy
\(S^{\mathfrak D}\).  Conformal covariance and graded permutation symmetry
are inherited from the covariance of the three-point tensor structures and
from item (vii).  Reflection positivity follows from the assumed positivity of
all inside-supported products; quotienting by zero-norm vectors gives the
physical radial Hilbert space.

The OPE of the reconstructed fields is built into the construction.  When a
subset of insertions is separated from the rest by a sphere, using an OPE
tree that first fuses that subset gives precisely the convergent expansion
with coefficients \(\lambda_{ab}^{c,\alpha}\) and conformally determined
descendant coefficients.  Since compatible trees give the same patched
correlator, the same OPE holds independently of the auxiliary tree used to
compute the correlator.  The stated uniqueness follows because any Euclidean
hierarchy with the same datum has the same convergent OPE sums on every
radial tree domain, and those domains cover separated configurations.

Finally, if tube-domain analytic continuation, the Lorentzian spectrum
condition, and local commutativity are included, the usual Wightman boundary
value construction applied to the reconstructed Euclidean hierarchy gives the
corresponding Lorentzian distributions.  These are precisely the additional
analytic hypotheses listed in
Hypothesis~\ref{hyp:radial-reconstruction-data}.
\end{proof}

\begin{openproblem}[Completeness of four-point OPE data]
\label{op:bootstrap-completeness-from-ope-data}
The useful reconstruction problem is not the claim that a bare crossing
equation for a single correlator produces a CFT, or even uniquely determines
that correlator.  The crossing equation is a functional constraint on an
unknown function or distribution with unknown spectral decomposition; without
additional input it has many formal solutions and many possible completions.
A single correlator probes only the sector generated by its external
operators, and even there one must resolve degeneracies, tensor structures,
normalizations, and null states.  The natural problem is sharper.  Suppose
one is given a complete all-primary OPE table
\[
  \{\Delta_a,\rho_a,\lambda_{ab}^{c,\alpha}\}
\]
or, in a generated theory, the complete OPE data appearing in the four-point
functions of a set of generators, with enough information to recover every
primary and every independent three-point tensor structure.  Suppose further
that this table obeys the unitary representation-theoretic bounds,
reflection-positivity inequalities, and the four-point associativity
equations for all relevant choices of external operators.  This is an exact
infinite-data problem, analogous to determining a theory from a sufficiently
complete set of precision observables; it is not the problem of solving a
single bootstrap equation.  Under what additional analytic hypotheses does
it extend to an abstract radial OPE datum satisfying
Definition~\ref{def:abstract-radial-ope-datum}?

In dimension \(D>2\), complete such data are expected to determine at most one
CFT up to unitary equivalence and changes of operator basis, once the
generator or all-primary hypothesis is made.  The existence part still
requires theorem-level control of radial convergence, spectral finiteness,
coherence of higher iterated OPEs, distributional extensions and contact-term
prescriptions, and Lorentzian boundary-value properties.  Those requirements
are precisely why Theorem~\ref{thm:conditional-cft-reconstruction-from-ope}
is formulated from an abstract radial OPE datum rather than from a single
informal crossing statement.
\end{openproblem}

\section{Conformal Blocks as Multiplet Contributions}

For a scalar four-point function, the OPE separates the contribution of each
primary conformal multiplet.  For the general scalar four-point function
\eqref{eq:scalar-four-point-general}, the \(12\to34\) OPE gives
\begin{equation}
  \mathcal G(u,v)
  =
  \sum_{\mathcal O,a,b}
  \lambda_{12\mathcal O}^{(a)}
  N_{\mathcal O}^{ab}
  \lambda_{34\mathcal O}^{(b)}
  G_{\Delta,\ell}^{(a,b;12|34)}(u,v).
  \label{eq:general-scalar-conformal-block-expansion}
\end{equation}
Here \(\mathcal O\) runs over primary multiplets of dimension \(\Delta\) and
spin \(\ell\), \(a,b\) label possible tensor structures in the
\(\mathcal O_1\mathcal O_2\mathcal O\) and
\(\mathcal O_3\mathcal O_4\mathcal O\) three-point functions, and
\(N_{\mathcal O}^{ab}\) is the inverse two-point pairing on the corresponding
degeneracy space.  In an orthonormal basis and for scalar external operators
with a single structure, this reduces to
\[
  \mathcal G(u,v)
  =
  \sum_{\mathcal O}
  \lambda_{12\mathcal O}\lambda_{34\mathcal O}\,
  G_{\Delta,\ell}^{(12|34)}(u,v).
\]
The function \(G_{\Delta,\ell}^{(12|34)}\) is the conformal block: it is the
sum of the primary \(\mathcal O\) and all of its descendants, with descendant
coefficients fixed by the conformal algebra.

\begin{figure}[H]
  \centering
  \resizebox{0.92\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      insertion/.style={circle,fill=violet!80!black,inner sep=1.7pt},
      line/.style={line width=0.8pt},
      arrow/.style={-{Latex[length=2mm]},line width=0.8pt}
    }
    \begin{scope}
      \node[font=\small] at (0,2.05) {\(12\to34\) channel};
      \node[insertion] at (-2.2,1.0) {};
      \node[insertion] at (-2.2,-1.0) {};
      \node[insertion] at (2.2,1.0) {};
      \node[insertion] at (2.2,-1.0) {};
      \node[violet!80!black,left] at (-2.2,1.0) {\(\mathcal O_1\)};
      \node[violet!80!black,left] at (-2.2,-1.0) {\(\mathcal O_2\)};
      \node[violet!80!black,right] at (2.2,1.0) {\(\mathcal O_4\)};
      \node[violet!80!black,right] at (2.2,-1.0) {\(\mathcal O_3\)};
      \draw[line] (-2.2,1.0)--(-0.75,0);
      \draw[line] (-2.2,-1.0)--(-0.75,0);
      \draw[line,blue!65!black] (-0.75,0)--(0.75,0);
      \draw[line] (0.75,0)--(2.2,1.0);
      \draw[line] (0.75,0)--(2.2,-1.0);
      \node[blue!65!black,above] at (0,0) {\(\mathcal O\)};
      \node[orange!85!black] at (-1.12,-0.34) {\(\lambda_{12\mathcal O}\)};
      \node[orange!85!black] at (1.10,-0.34) {\(\lambda_{34\mathcal O}\)};
    \end{scope}
    \draw[arrow] (3.35,0)--(4.65,0);
    \node at (4.00,0.48) {same correlator};
    \begin{scope}[shift={(7.05,0)}]
      \node[font=\small] at (0,2.05) {\(14\to23\) channel};
      \node[insertion] at (-2.2,1.0) {};
      \node[insertion] at (-2.2,-1.0) {};
      \node[insertion] at (2.2,1.0) {};
      \node[insertion] at (2.2,-1.0) {};
      \node[violet!80!black,left] at (-2.2,1.0) {\(\mathcal O_1\)};
      \node[violet!80!black,left] at (-2.2,-1.0) {\(\mathcal O_2\)};
      \node[violet!80!black,right] at (2.2,1.0) {\(\mathcal O_4\)};
      \node[violet!80!black,right] at (2.2,-1.0) {\(\mathcal O_3\)};
      \draw[line] (-2.2,1.0)--(-0.20,0.72);
      \draw[line] (2.2,1.0)--(-0.20,0.72);
      \draw[line,blue!65!black] (-0.20,0.72)--(0.20,-0.72);
      \draw[line] (-2.2,-1.0)--(0.20,-0.72);
      \draw[line] (2.2,-1.0)--(0.20,-0.72);
      \node[blue!65!black,right] at (0.18,0) {\(\mathcal O'\)};
      \node[orange!85!black] at (-0.65,1.10) {\(\lambda_{14\mathcal O'}\)};
      \node[orange!85!black] at (-0.55,-1.16) {\(\lambda_{23\mathcal O'}\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{A four-point function may be expanded in different OPE trees.  The
  equality of these decompositions is the four-point form of OPE
  associativity.}
  \label{fig:ope-crossing-trees}
\end{figure}

The block is characterized representation-theoretically by the quadratic
Casimir, together with an OPE boundary condition.  Put
\[
  u=z\bar z,\qquad v=(1-z)(1-\bar z),
  \qquad
  a=\frac{\Delta_2-\Delta_1}{2},\qquad
  b=\frac{\Delta_3-\Delta_4}{2},
  \qquad
  \varepsilon_D=\frac D2-1 .
\]
After the external scalar prefactor in
\eqref{eq:scalar-four-point-general} has been removed, the \(12\)-channel
quadratic Casimir acts on the reduced function by
\begin{equation}
  \mathcal C_{12}^{(u,v)}=2\mathcal D_{a,b},
  \label{eq:scalar-casimir-uv-definition}
\end{equation}
where
\begin{align}
  \mathcal D_{a,b}
  &=
  u^2(1-u+v)\partial_u^2
  -2uv(1+u-v)\partial_u\partial_v
  +v\bigl((1-v)^2-u(1+v)\bigr)\partial_v^2
  \notag\\
  &\quad
  -u\bigl((a+b+1)(1+u-v)+2\varepsilon_D\bigr)\partial_u
  +(a+b+1)\bigl((1-v)^2-u(1+v)\bigr)\partial_v
  \notag\\
  &\quad
  -ab(1+u-v).
  \label{eq:scalar-casimir-uv}
\end{align}
Equivalently, in the algebraic coordinates \(z,\bar z\),
\begin{align}
  \mathcal D_{a,b}
  &=
  D_z+D_{\bar z}
  +
  2\varepsilon_D
  \frac{z\bar z}{z-\bar z}
  \bigl((1-z)\partial_z-(1-\bar z)\partial_{\bar z}\bigr),
  \label{eq:scalar-casimir-zbarz}
  \\
  D_z
  &=
  z^2(1-z)\partial_z^2
  -(a+b+1)z^2\partial_z
  -abz,
  \qquad
  D_{\bar z}=D_z\big|_{z\mapsto\bar z}.
  \notag
\end{align}
The conversion from \eqref{eq:scalar-casimir-zbarz} to
\eqref{eq:scalar-casimir-uv} is the chain rule together with
\[
  z+\bar z=1+u-v,\qquad
  \partial_z=\bar z\,\partial_u-(1-\bar z)\partial_v,\qquad
  \partial_{\bar z}=z\,\partial_u-(1-z)\partial_v .
\]
The apparently singular term in \eqref{eq:scalar-casimir-zbarz} is regular on
symmetric functions of \(z,\bar z\), since
\[
  (1-z)\partial_z-(1-\bar z)\partial_{\bar z}
  =
  (\bar z-z)\partial_u .
\]
The factor \(2\) in \eqref{eq:scalar-casimir-uv-definition} fixes the
normalization of the conformal-algebra quadratic Casimir used in this chapter:
\begin{equation}
  \mathcal C_{12}^{(u,v)}
  G_{\Delta,\ell}^{(12|34)}(u,v)
  =
  C_{\Delta,\ell}
  G_{\Delta,\ell}^{(12|34)}(u,v),
  \qquad
  C_{\Delta,\ell}
  =
  \Delta(\Delta-D)+\ell(\ell+D-2).
  \label{eq:scalar-block-casimir-equation}
\end{equation}
The equality \(C_{\Delta,\ell}=C_{D-\Delta,\ell}\) already shows that the
differential equation alone does not select an OPE block.  In the radial frame
of \eqref{eq:rho-cross-ratio-map}, the two local Frobenius branches at the
\(12\) OPE boundary are
\begin{align}
  G_{\Delta,\ell}^{(12|34)}(u,v)
  &=
  r^\Delta C_\ell^{(D/2-1)}(\cos\alpha)
  +O(r^{\Delta+1}),
  \label{eq:block-ope-frobenius-branch}
  \\
  G_{D-\Delta,\ell}^{(12|34)}(u,v)
  &=
  r^{D-\Delta} C_\ell^{(D/2-1)}(\cos\alpha)
  +O(r^{D-\Delta+1}).
  \label{eq:block-shadow-frobenius-branch}
\end{align}
The conformal block in the OPE sum is the solution with the first leading
coefficient normalized as in
\eqref{eq:block-ope-frobenius-branch} and with zero coefficient multiplying the
shadow branch \eqref{eq:block-shadow-frobenius-branch}.

Short multiplets require a levelwise quotient, not a single informal deletion
of dependent vectors.  Section~\ref{sec:short-multiplets-operator-equations}
explains how saturation of a unitarity bound produces a null descendant at a
definite level: for example the scalar Laplace descendant
\(\widehat P^2\ket\phi\) at level \(2\), the divergence of a conserved
symmetric-traceless tensor at level \(1\), or the spinor gamma-trace descendant
at level \(1\).  If \(N_{\alpha}\) denotes such a singular null descendant at
level \(r_\alpha\), then every
\[
  \widehat P_{\mu_1}\cdots\widehat P_{\mu_k}N_{\alpha}
  \qquad (k\ge0)
\]
is null at level \(r_\alpha+k\).  The descendant space used in the block at
level \(n\) is therefore the raw level-\(n\) descendant space modulo the span of
all null descendants whose total level is \(n\).  This is the same null
submodule subtraction that produces the short-module characters in
Chapter~\ref{ch:volume-iii-radial-quantization}; it is performed before any
Gram matrix is inverted.

\begin{theorem}[Local existence and uniqueness of the OPE block]
\label{thm:casimir-block-local-existence-uniqueness}
Assume the radial reconstruction hypotheses of
Hypothesis~\ref{hyp:radial-reconstruction-data}, finite multiplicity of
descendant states at each level, and the levelwise null-submodule quotient just
described.  For scalar external operators and one exchanged symmetric-traceless
primary of dimension \(\Delta\) and spin \(\ell\), there is a unique
\(12\)-channel block with the OPE normalization
\eqref{eq:block-ope-frobenius-branch}.  It is a convergent radial series for
\(r<1\), satisfies the Casimir equation
\eqref{eq:scalar-block-casimir-equation}, and contains no shadow branch.
\end{theorem}

\begin{proof}
Let \(V_{\Delta,\ell}\) be the conformal multiplet generated from the exchanged
primary.  Let
\[
  W_n\simeq \operatorname{Sym}^n(\mathbb C^D)\otimes V_{\Delta,\ell}
\]
be the raw level-\(n\) descendant space before equations of motion,
conservation laws, or other shortening relations are imposed.  Its null radical
for the radial Gram form is
\[
  \mathcal N_n
  =
  \{v\in W_n:\langle v,w\rangle=0\text{ for all }w\in W_n\}.
\]
When the exchanged multiplet is short, \(\mathcal N_n\) is generated
level-by-level by the singular null descendants fixed by the saturated
unitarity bounds and by all of their translation descendants at total level
\(n\).  Reflection positivity implies that every vector in \(\mathcal N_n\) has
zero matrix element against every physical exterior state, so replacing
\(W_n\) by
\[
  Q_n=W_n/\mathcal N_n
\]
does not change any physical four-point function.  Choose a basis
\(\{\ket{A,n}\}_A\) of \(Q_n\) and let
\[
  M^{(n)}_{AB}=\langle A,n|B,n\rangle
\]
be the quotient Gram matrix.  The quotient has no null radical, hence
\(M^{(n)}\) is invertible on this finite-dimensional space.  The contribution
of this level to the four-point function is
\begin{equation}
  r^{\Delta+n}
  \sum_{A,B}
  \langle \mathcal O_4\mathcal O_3|A,n\rangle\,
  (M^{(n)})^{-1}_{AB}\,
  \langle B,n|\mathcal O_1\mathcal O_2\rangle ,
  \label{eq:block-level-n-gram-construction}
\end{equation}
with the angular dependence carried by the cylinder wavefunctions of the
descendants.  The three-point Ward identities fix the two vectors of matrix
elements in \eqref{eq:block-level-n-gram-construction} recursively from the
primary three-point structures, because moving a conformal generator from a
descendant state onto the external primaries is governed by the conformal
algebra and by the scalar transformation law.  Summing over all \(n\geq0\)
therefore constructs a radial series whose leading term is the primary
matrix element
\[
  r^\Delta C_\ell^{(D/2-1)}(\cos\alpha).
\]
The convergence for \(r<1\) is the radial OPE convergence theorem applied after
projecting the inside state to the multiplet \(V_{\Delta,\ell}\).

The same construction gives the Casimir equation.  The operator
\(\mathcal C_{12}\) is the square of the total conformal generator acting on the
pair \((1,2)\).  On the projected intermediate multiplet it acts by the scalar
\(C_{\Delta,\ell}\), while the external scalar prefactor converts this action
on the four-point function into the differential operator
\(\mathcal C_{12}^{(u,v)}\) displayed in
\eqref{eq:scalar-casimir-uv}.  Hence the constructed radial series satisfies
\eqref{eq:scalar-block-casimir-equation}.

For uniqueness, let \(H(r,\alpha)\) be the difference of two local solutions
with the OPE normalization \eqref{eq:block-ope-frobenius-branch} and no shadow
branch.  Its first nonzero radial coefficient, if present, has some level
\(n>0\) and lies in the finite descendant space at that level.  The Casimir
equation and the Ward identities impose the same finite linear system as in
\eqref{eq:block-level-n-gram-construction}.  Since \(M^{(n)}\) is invertible on
the quotient by null states and all lower-level coefficients of \(H\) vanish,
that coefficient must vanish.  This contradicts the choice of the first
nonzero level.  Therefore \(H=0\), proving uniqueness.
\end{proof}

Thus the spectrum and OPE coefficients are the dynamical data, while the
blocks are kinematic functions determined by conformal representation theory
and by the OPE branch condition.

\subsection{Blocks, Shadows, and Partial Waves}

The object just defined is an OPE-channel conformal block.  Its normalization
in this chapter is the OPE normalization: in the \(12\to34\) channel the
leading term as \(x_1\to x_2\) is the contribution of a primary of dimension
\(\Delta\) and spin \(\ell\), with no contribution from the shadow dimension.
This is the normalization used in the discrete OPE sum
\eqref{eq:general-scalar-conformal-block-expansion}.

A \emph{conformal partial wave} is a different, although closely related,
object.  It is the harmonic-analysis kernel obtained from the conformal
projector onto an irreducible principal-series representation.  Let
\(\widetilde\Delta=D-\Delta\) be the shadow dimension of
\(\Delta\).  For scalar external operators and a single exchanged
symmetric-traceless tensor structure, the shadow-projector normalization may
be represented, with the symmetric-traceless spin-index contraction suppressed
from the notation, by
\begin{equation}
  \Psi_{\Delta,\ell}^{(12|34)}(x_i)
  =
  \frac{1}{\mathcal N_{\Delta,\ell}}
  \int_{\R^D}\dd^D x_0\,
  \bigl\langle
    \mathcal O_1(x_1)\mathcal O_2(x_2)
    \mathcal O_{\Delta,\ell}(x_0)
  \bigr\rangle
  \bigl\langle
    \widetilde{\mathcal O}_{\widetilde\Delta,\ell}(x_0)
    \mathcal O_3(x_3)\mathcal O_4(x_4)
  \bigr\rangle .
  \label{eq:shadow-projector-partial-wave}
\end{equation}
The spin indices of
\(\mathcal O_{\Delta,\ell}\) and
\(\widetilde{\mathcal O}_{\widetilde\Delta,\ell}\) are contracted with the
invariant two-point pairing.  The constant
\(\mathcal N_{\Delta,\ell}\) is fixed once the two-point functions and the
shadow transform are normalized; the convention is that the resulting
projector is idempotent on the irreducible representation and that the
Plancherel formula below has no additional scalar normalization.  With
spinning external operators or multiple three-point tensor structures,
\(\Psi_{\Delta,\ell}\) becomes a matrix in the corresponding structure labels.

After the same external kinematic prefactor has been factored out as in the
four-point decomposition, the partial wave gives a function
\(\Psi_{\Delta,\ell}^{(12|34)}(u,v)\).  It satisfies the same quadratic
Casimir equation as the block with eigenvalue \(C_{\Delta,\ell}\), but it has
different analytic boundary data.  In the \(12\) OPE limit it is a fixed
linear combination
\begin{equation}
  \Psi_{\Delta,\ell}^{(12|34)}(u,v)
  =
  A_{\Delta,\ell}^{(12|34)}
  G_{\Delta,\ell}^{(12|34)}(u,v)
  +
  A_{\widetilde\Delta,\ell}^{(12|34)}
  G_{\widetilde\Delta,\ell}^{(12|34)}(u,v),
  \qquad
  \widetilde\Delta=D-\Delta ,
  \label{eq:partial-wave-block-shadow-block}
\end{equation}
where the coefficients \(A_{\Delta,\ell}\) are determined by the chosen
shadow-transform normalization.  Thus a partial wave is not the same object
as the conformal block appearing in a physical Euclidean OPE sum.  The block
selects one OPE boundary condition; the partial wave is a single-valued
Euclidean conformal harmonic, containing both the block and the shadow block.

For functions of cross-ratios satisfying the hypotheses of conformal harmonic
analysis, the partial waves on the principal series
\(\Delta=D/2+\ii\nu\), \(\nu\in\R\), form the continuous basis
\begin{equation}
  \mathcal F(u,v)
  =
  \sum_{\ell=0}^{\infty}
  \int_{-\infty}^{\infty}\frac{\dd\nu}{2\pi}\,
  \mu(\nu,\ell)\,
  a(\nu,\ell)\,
  \Psi_{D/2+\ii\nu,\ell}(u,v).
  \label{eq:conformal-partial-wave-plancherel}
\end{equation}
Here \(\mu(\nu,\ell)\) is the Plancherel density associated with the
normalization in \eqref{eq:shadow-projector-partial-wave}, and
\(a(\nu,\ell)\) is the partial-wave transform of \(\mathcal F\).  Physical
OPE data are recovered, when the required analyticity and growth hypotheses
hold, by moving the \(\nu\)-contour and taking residues at poles in
\(\Delta\).  This is why one must distinguish the two objects: conformal
blocks organize the discrete OPE in a channel, whereas conformal partial
waves provide the representation-theoretic resolution of identity used in
integral inversion formulae.

\subsection{Mellin Representation and OPE Poles}

Mellin space is a transform representation of scalar separated-point
correlators.  Its use requires an existence and growth hypothesis parallel to
the partial-wave transform above: a meromorphic Mellin amplitude belongs to a
specific Mellin-Barnes representation of the correlator.  When the
representation exists with suitable vertical-contour decay, it turns the
\(12\)-channel OPE expansion into a residue expansion in a complex variable
conjugate to the cross-ratio \(u\).

Let \(X_{ij}=x_{ij}^2\) for Euclidean separated points and let
\(\mathcal O_i\) be scalar primaries of dimensions \(\Delta_i\).  Define the
affine Mellin constraint space
\begin{equation}
  \mathfrak A_{\Delta}
  =
  \left\{
    \delta_{ij}=\delta_{ji}\in\mathbb C,\quad
    \delta_{ii}=0,\quad
    \sum_{j\ne i}\delta_{ij}=\Delta_i
    \text{ for }i=1,\ldots,n
  \right\}.
  \label{eq:mellin-affine-constraint-space}
\end{equation}
The constraints are exactly the conditions that make
\(\prod_{i<j}X_{ij}^{-\delta_{ij}}\) transform like an \(n\)-point function
of scalar primaries under independent local scale factors at the insertion
points.

\begin{definition}[Scalar Mellin representation]
\label{def:scalar-mellin-representation}
An \(n\)-point scalar Euclidean correlator is Mellin representable on a
configuration-space domain \(U\) if there are vertical cycles
\(\mathcal C\subset\mathfrak A_\Delta\), a meromorphic or distributional
function \(M(\delta_{ij})\), and a choice of logarithm for each \(X_{ij}\) on
\(U\), such that
\begin{equation}
  \left\langle
    \mathcal O_1(x_1)\cdots\mathcal O_n(x_n)
  \right\rangle
  =
  \int_{\mathcal C}[\dd\delta]\,
  M(\delta_{ij})
  \prod_{1\le i<j\le n}
  \Gamma(\delta_{ij})\,X_{ij}^{-\delta_{ij}} .
  \label{eq:mellin-representation-n-point-core}
\end{equation}
The measure \([\dd\delta]\) is the product of \(\dd\delta/(2\pi\ii)\) over a
basis of independent affine coordinates on
\(\mathfrak A_\Delta\).  The contour is chosen so that the pole families whose
residues give the desired OPE expansions lie on the side prescribed by the
corresponding radial domain.  The gamma factors in
\eqref{eq:mellin-representation-n-point-core} are part of the normalization
of \(M\); moving any universal gamma factor between \(M\) and the kernel gives
a different Mellin amplitude representing the same correlator.
\end{definition}

For four points it is useful to replace the constrained variables by three
Mellin-channel variables
\[
  \mathfrak s+\mathfrak t+\mathfrak u
  =
  \Delta_\Sigma,
  \qquad
  \Delta_\Sigma=\Delta_1+\Delta_2+\Delta_3+\Delta_4,
\]
where
\begin{align}
  \delta_{12}&=\frac{\Delta_1+\Delta_2-\mathfrak s}{2},
  &
  \delta_{34}&=\frac{\Delta_3+\Delta_4-\mathfrak s}{2},
  \notag\\
  \delta_{14}&=\frac{\Delta_1+\Delta_4-\mathfrak t}{2},
  &
  \delta_{23}&=\frac{\Delta_2+\Delta_3-\mathfrak t}{2},
  \label{eq:mellin-four-point-delta-variables}\\
  \delta_{13}&=\frac{\Delta_1+\Delta_3-\mathfrak u}{2},
  &
  \delta_{24}&=\frac{\Delta_2+\Delta_4-\mathfrak u}{2}.
  \notag
\end{align}
The letters \(\mathfrak s,\mathfrak t,\mathfrak u\) are Mellin variables;
they are not scattering Mandelstam invariants.  They are named by the OPE
channel whose twist poles they expose.

With the four-point prefactor of
\eqref{eq:scalar-four-point-general}, Definition
\ref{def:scalar-mellin-representation} becomes
\begin{equation}
  \mathcal G(u,v)
  =
  \int_{\mathcal C_{\mathfrak s,\mathfrak t}}
  \frac{\dd\mathfrak s\,\dd\mathfrak t}{(4\pi\ii)^2}\,
  u^{\mathfrak s/2}
  v^{(\mathfrak t-\Delta_2-\Delta_3)/2}
  \Gamma_\Delta(\mathfrak s,\mathfrak t)\,
  M(\mathfrak s,\mathfrak t),
  \label{eq:four-point-mellin-representation-core}
\end{equation}
where \(\mathfrak u=\Delta_\Sigma-\mathfrak s-\mathfrak t\) and
\begin{align}
  \Gamma_\Delta(\mathfrak s,\mathfrak t)
  &=
  \Gamma\!\left(\frac{\Delta_1+\Delta_2-\mathfrak s}{2}\right)
  \Gamma\!\left(\frac{\Delta_3+\Delta_4-\mathfrak s}{2}\right)
  \Gamma\!\left(\frac{\Delta_1+\Delta_4-\mathfrak t}{2}\right)
  \Gamma\!\left(\frac{\Delta_2+\Delta_3-\mathfrak t}{2}\right)
  \notag\\
  &\quad\times
  \Gamma\!\left(\frac{\Delta_1+\Delta_3-\mathfrak u}{2}\right)
  \Gamma\!\left(\frac{\Delta_2+\Delta_4-\mathfrak u}{2}\right).
  \label{eq:four-point-mellin-gamma-kernel}
\end{align}
The normalization \((4\pi\ii)^{-2}\) comes from
\(\dd\delta_{12}\dd\delta_{14}=\dd\mathfrak s\,\dd\mathfrak t/4\).

\begin{lemma}[Compatibility with the four-point prefactor]
\label{lem:mellin-four-point-prefactor-compatibility}
The representation \eqref{eq:four-point-mellin-representation-core} is exactly
the \(n=4\) constrained Mellin representation
\eqref{eq:mellin-representation-n-point-core} after factoring out the
kinematic prefactor in \eqref{eq:scalar-four-point-general}.
\end{lemma}

\begin{proof}
The Mellin product has exponent \(-\delta_{ij}\) on \(X_{ij}\).  Subtract the
exponents of the prefactor in \eqref{eq:scalar-four-point-general}.  For
\(X_{12}\) and \(X_{34}\) the remaining exponent is
\[
  -\delta_{12}+\frac{\Delta_1+\Delta_2}{2}
  =
  -\delta_{34}+\frac{\Delta_3+\Delta_4}{2}
  =
  \frac{\mathfrak s}{2}.
\]
For \(X_{14}\) and \(X_{23}\) the remaining exponent is
\[
  \frac{\mathfrak t-\Delta_2-\Delta_3}{2}.
\]
For \(X_{13}\) and \(X_{24}\) the remaining exponent is the negative of the
sum of these two numbers, because
\(\mathfrak u=\Delta_\Sigma-\mathfrak s-\mathfrak t\).  These are precisely
the exponents of
\[
  u^{\mathfrak s/2}v^{(\mathfrak t-\Delta_2-\Delta_3)/2}
  =
  \left(\frac{X_{12}X_{34}}{X_{13}X_{24}}\right)^{\mathfrak s/2}
  \left(\frac{X_{14}X_{23}}{X_{13}X_{24}}\right)^{
        (\mathfrak t-\Delta_2-\Delta_3)/2}.
 \]
The six gamma factors are those in
\eqref{eq:mellin-representation-n-point-core} with
\eqref{eq:mellin-four-point-delta-variables} substituted.
\end{proof}

The pole interpretation is a contour-shift statement.  Write
\[
  \mathcal I(\mathfrak s,\mathfrak t)
  =
  \Gamma_\Delta(\mathfrak s,\mathfrak t)M(\mathfrak s,\mathfrak t)
\]
for the full Mellin integrand without the cross-ratio powers.  Suppose the
\(\mathfrak s\)-contour is a vertical line, \(0<u<1\), the integrand is
meromorphic in the strip swept by moving the contour to the right, and the
horizontal contour arcs vanish in that strip.  If
\(\mathcal I\) has a simple pole at \(\mathfrak s=\sigma\) with residue
\(R_\sigma(\mathfrak t)\), then that pole contributes
\begin{equation}
  u^{\sigma/2}
  \int_{\mathcal C_{\mathfrak t}}
  \frac{\dd\mathfrak t}{2\pi\ii}\,
  v^{(\mathfrak t-\Delta_2-\Delta_3)/2}
  R_\sigma(\mathfrak t)
  \label{eq:mellin-simple-pole-u-exponent}
\end{equation}
to \(\mathcal G(u,v)\).  A pole of order \(r+1\) gives a polynomial of degree
\(r\) in \(\log u\) multiplying \(u^{\sigma/2}\).  Conversely, under the same
vertical-growth hypothesis, a term \(u^\alpha H(v)\) in the asymptotic
expansion at fixed \(v\) is represented in the Mellin transform by a pole at
\(\mathfrak s=2\alpha\), with residue equal to the one-dimensional Mellin
transform of \(H(v)\) in the \(\mathfrak t\) variable.

Now apply this to the radial OPE.  An exchanged primary of dimension
\(\Delta\) and spin \(\ell\) has twist
\[
  \tau=\Delta-\ell .
\]
In the \(12\to34\) OPE channel, the corresponding conformal block has a
radial expansion of the form
\begin{equation}
  G_{\Delta,\ell}^{(12|34)}(u,v)
  =
  u^{\tau/2}
  \sum_{m=0}^{\infty}u^m\,H_m^{(\Delta,\ell)}(v),
  \label{eq:block-small-u-twist-expansion}
\end{equation}
inside the radial convergence domain, after the spin-dependent angular
dependence is expressed through \(v\) in the chosen conformal frame.  Therefore
the full Mellin integrand has poles at
\begin{equation}
  \mathfrak s=\tau+2m,\qquad m=0,1,2,\ldots .
  \label{eq:mellin-ope-twist-pole-family-core}
\end{equation}
If the gamma-stripped amplitude \(M\) is meromorphic and the pole does not
coincide with a gamma-kernel pole, the same family appears in \(M\).  With the
OPE normalization of the block fixed in
\eqref{eq:block-ope-frobenius-branch}, its local form is
\begin{equation}
  M(\mathfrak s,\mathfrak t)
  =
  \sum_{m=0}^{\infty}
  \frac{
    \lambda_{12\mathcal O}\lambda_{34\mathcal O}\,
    Q_{m,\ell}^{(12|34)}(\mathfrak t)
  }{\mathfrak s-\tau-2m}
  +
  M_{\rm hol}(\mathfrak s,\mathfrak t),
  \label{eq:mellin-mack-pole-family}
\end{equation}
in a neighborhood of the displayed pole family.  The residue
\(Q_{m,\ell}^{(12|34)}\) is the Mack polynomial in this normalization.  It is
characterized intrinsically as the unique polynomial residue which, after
substitution in \eqref{eq:four-point-mellin-representation-core}, produces
the level-\(m\) term of the Casimir eigenfunction with OPE boundary condition
\eqref{eq:block-ope-frobenius-branch}.  Its degree in \(\mathfrak t\) is
\(\ell\); the descendant label \(m\) changes the coefficients and the gamma
kernel, not the spin degree.

\begin{proposition}[OPE poles in Mellin space]
\label{prop:mellin-ope-pole-family}
Assume the radial OPE theorem of
Theorem~\ref{thm:euclidean-radial-ope-convergence}, the four-point Mellin
representation \eqref{eq:four-point-mellin-representation-core}, meromorphic
continuation of \(M\) in the relevant \(\mathfrak s\)-strip, and the vertical
growth needed for the contour shift above.  Then every exchanged
symmetric-traceless primary \(\mathcal O\) in the \(12\to34\) scalar OPE whose
twist-pole family does not coincide with a gamma-kernel pole contributes the
pole family \eqref{eq:mellin-mack-pole-family}.  The residue is
proportional to the product
\(\lambda_{12\mathcal O}\lambda_{34\mathcal O}\), and the proportionality
function is fixed by conformal representation theory.
\end{proposition}

\begin{proof}
The radial OPE theorem gives an absolutely convergent sum of conformal
families in the channel domain.  Project to a single exchanged conformal
multiplet and use the block construction in
Theorem~\ref{thm:casimir-block-local-existence-uniqueness}.  Its radial
series has the form
\eqref{eq:block-small-u-twist-expansion}: the primary supplies the exponent
\(\tau/2\), and the descendant contribution can be grouped into integer powers
of \(u\) after the remaining angular dependence is kept in the functions
\(H_m(v)\).  The coefficient multiplying the block in the four-point function
is
\(\lambda_{12\mathcal O}\lambda_{34\mathcal O}\) in the orthonormal
single-structure case, or the corresponding contraction with the inverse
two-point pairing in the notation of
\eqref{eq:general-scalar-conformal-block-expansion}.

The contour-shift identity
\eqref{eq:mellin-simple-pole-u-exponent} identifies every term
\(u^{\tau/2+m}H_m(v)\) with a pole at
\(\mathfrak s=\tau+2m\) in the full Mellin integrand.  When this pole is not a
pole already forced by \(\Gamma_\Delta\), it is represented by a pole of
\(M\).  The \(v\)-dependence of the residue is converted to a polynomial in
\(\mathfrak t\) by the one-dimensional Mellin transform with the fixed gamma
kernel in \eqref{eq:four-point-mellin-gamma-kernel}.  The Casimir equation
\eqref{eq:scalar-block-casimir-equation} and the OPE boundary condition
\eqref{eq:block-ope-frobenius-branch} fix this residue uniquely.  The highest
power of \(\mathfrak t\) is fixed by the spin-\(\ell\) leading angular
structure of the primary three-point functions, hence the residue has degree
\(\ell\).  This proves the stated pole family and normalization.
\end{proof}

Crossing acts linearly on Mellin space.  A permutation \(\sigma\in S_4\) sends
\(\delta_{ij}\) to \(\delta_{\sigma(i)\sigma(j)}\), and the induced action on
\((\mathfrak s,\mathfrak t,\mathfrak u)\) permutes the channel variables with
\(\mathfrak s+\mathfrak t+\mathfrak u=\Delta_\Sigma\) preserved.  For
identical scalar primaries, crossing of the separated-point correlator is
equivalent, in a common Mellin domain, to the corresponding invariance of the
Mellin integrand up to contour deformation across the pole families.  Thus the
Mellin form expresses the same OPE-associativity constraints as the block
expansion, but in a language where exchanged twists are pole locations and
spins are polynomial degrees of residues.

\subsection{Lorentzian Inversion as a Boundary Theorem}

The Euclidean OPE sum determines correlators by summing over the discrete
operator spectrum.  Lorentzian inversion goes in the opposite direction: under
additional Lorentzian analyticity and boundedness hypotheses, it reconstructs
the spin-analytic OPE coefficient function from a double discontinuity of the
four-point function.  For identical scalar primaries, define the \(t\)-channel
double discontinuity across the branch cut \(\bar z\in[1,\infty)\) by
\begin{equation}
  \operatorname{dDisc}_t[\mathcal G](z,\bar z)
  =
  \mathcal G(z,\bar z)
  -
  \frac12\,\mathcal G^{\circlearrowleft}(z,\bar z)
  -
  \frac12\,\mathcal G^{\circlearrowright}(z,\bar z),
  \label{eq:lorentzian-ddisc-identical}
\end{equation}
where the arrows denote the two analytic continuations of \(\bar z\) around
\(\bar z=1\).  For unequal external dimensions the two continued terms are
multiplied by the corresponding monodromy phases fixed by the external
weights.

In the scalar normalization compatible with the partial-wave convention above,
the \(t\)-channel contribution to the Caron-Huot Lorentzian inversion formula
has the form
\begin{equation}
  c^t(\Delta,J)
  =
  \frac{\kappa_{\Delta+J}}{4}
  \int_0^1\dd z\int_0^1\dd\bar z\,
  \frac{|z-\bar z|^{D-2}}{(z\bar z)^D}\,
  G_{J+D-1,\,\Delta-D+1}^{(12|34)}(z,\bar z)\,
  \operatorname{dDisc}_t[\mathcal G](z,\bar z),
  \label{eq:caron-huot-lorentzian-inversion-identical}
\end{equation}
with
\begin{equation}
  \kappa_\beta
  =
  \frac{\Gamma(\beta/2)^4}
       {2\pi^2\,\Gamma(\beta-1)\Gamma(\beta)}
  \qquad
  \text{for identical external scalars.}
  \label{eq:caron-huot-kappa-identical}
\end{equation}
The full identical-scalar coefficient includes the crossed \(u\)-channel
contribution
\[
  c(\Delta,J)=c^t(\Delta,J)+(-1)^Jc^u(\Delta,J),
\]
with the same kernel after the corresponding interchange of points.  The block
appearing in the kernel has analytically continued labels: its first label is
\(J+D-1\), and its second label is \(\Delta-D+1\).  This is the same
block/partial-wave duality already visible in
\eqref{eq:partial-wave-block-shadow-block}; the inversion integral pairs the
Lorentzian double discontinuity with the shadow-side harmonic kernel.

The coefficient function \(c(\Delta,J)\) is meromorphic in \(\Delta\) for the
spins covered by the Regge boundedness hypothesis.  If a primary of spin \(J\)
and dimension \(\Delta_{\mathcal O}\) appears in the \(\phi\times\phi\) OPE,
then near the pole
\begin{equation}
  c(\Delta,J)
  =
  \frac{\lambda_{\phi\phi\mathcal O}^{\,2}}
       {\Delta-\Delta_{\mathcal O}}
  +\text{holomorphic near }\Delta_{\mathcal O}.
  \label{eq:lorentzian-inversion-pole-extraction}
\end{equation}
Thus the formula extracts OPE data from Lorentzian branch-cut data rather than
from Euclidean projection onto a discrete radial spectrum.

\begin{remark}[Hypotheses and scope]
\label{rem:lorentzian-inversion-scope}
Equation~\eqref{eq:caron-huot-lorentzian-inversion-identical} is not an
additional axiom of CFT data.  It is a theorem under hypotheses: unitarity or
reflection positivity sufficient for the Lorentzian continuation, Euclidean OPE
convergence in the required domains, analyticity in the relevant tube domains,
and Regge boundedness strong enough to suppress the arcs used in the contour
deformation.  The spin range begins above the Regge intercept; in many unitary
CFT applications this gives \(J>1\), with possible lower-spin terms requiring
separate subtractions.  Later analytic-bootstrap and light-ray discussions use
the inversion formula only at this theorem boundary.\footnote{See
S.~Caron-Huot, ``Analyticity in spin in conformal theories,'' JHEP
\textbf{09} (2017) 078, arXiv:1703.00278; D.~Simmons-Duffin,
D.~Stanford, and E.~Witten, ``A spacetime derivation of the Lorentzian OPE
inversion formula,'' JHEP \textbf{07} (2018) 085, arXiv:1711.03816; and
P.~Kravchuk and D.~Simmons-Duffin, ``Light-ray operators in conformal field
theory,'' JHEP \textbf{11} (2018) 102, arXiv:1805.00098.}
\end{remark}

\subsection{Radial Expansion of a Block}

For identical Hermitian scalar primaries \(\phi\), there is a particularly
transparent block expansion in the Hogervorst--Rychkov radial variable
\(\rho\).  Put the four points in a plane at
\[
  x_1=\rho,\qquad x_2=-\rho,\qquad x_3=-1,\qquad x_4=1,
  \qquad
  \rho=r\ee^{\ii\alpha},\quad 0\le r<1.
\]
With the cross-ratio convention of
\eqref{eq:cross-ratios-u-v}, this frame gives
\begin{equation}
  z=\frac{4\rho}{(1+\rho)^2},
  \qquad
  \rho=\frac{z}{(1+\sqrt{1-z})^2}.
  \label{eq:rho-cross-ratio-map}
\end{equation}
The Weyl map to the cylinder places the two inner insertions at radial time
\(\tau=\log r\) and angular positions separated by \(\pi\), while the two
outer insertions lie at \(\tau=0\).  The map
\eqref{eq:rho-cross-ratio-map} is chosen so that \(|\rho|<1\) is precisely the
radial ordering domain for the \(12\to34\) OPE channel; the descendant expansion
below is the radial-coordinate expansion introduced by Hogervorst and
Rychkov.\footnote{See M.~Hogervorst and S.~Rychkov, ``Radial Coordinates for
Conformal Blocks,'' Phys. Rev. D \textbf{87} (2013) 106004,
arXiv:1303.1111.}

\begin{theorem}[Hogervorst--Rychkov radial block convergence]
\label{thm:hogervorst-rychkov-rho-convergence}
Assume the radial reconstruction hypotheses of
Hypothesis~\ref{hyp:radial-reconstruction-data}, finite multiplicity of radial
energy eigenspaces, and reflection positivity in the channel under
consideration.  For each exchanged conformal family in a scalar four-point
function, the radial block expansion constructed below converges absolutely,
with convergence uniform on every closed subdisk \(|\rho|\le q<1\), in the
Euclidean radial configuration where \(\bar\rho=\rho^*\).  Equivalently, after
analytic continuation of \(z\) with the square-root branch
\(\operatorname{Re}\sqrt{1-z}>0\),
\[
  \rho(z)
  =
  \frac{z}{(1+\sqrt{1-z})^2}
  =
  \frac{1-\sqrt{1-z}}{1+\sqrt{1-z}}
\]
maps the cut \(z\)-plane \(\mathbb C\setminus[1,\infty)\) to the unit disk,
and every compact subset of the cut plane maps into \(|\rho|\le q\) for some
\(q<1\).  The radial series therefore gives the analytic continuation of the
block throughout that regularity domain.
\end{theorem}

\begin{proof}
The geometric part is elementary.  Put \(w=\sqrt{1-z}\) with
\(\operatorname{Re}w>0\).  Then
\[
  \rho=\frac{1-w}{1+w}.
\]
The Cayley transform sends the right half-plane to the unit disk because
\[
  |\rho|^2
  =
  \frac{|1-w|^2}{|1+w|^2}
  =
  1-\frac{4\operatorname{Re}w}{|1+w|^2}<1 .
\]
Its inverse is \(z=4\rho/(1+\rho)^2\).  Hence the unit disk in \(\rho\) is not
only the old \(|z|<1\) power-series domain in disguise; it is the
uniformizing radial disk for the whole \(z\)-plane cut along the crossed
channel singular locus.

The analytic convergence is the radial Hilbert-space argument in the symmetric
cylinder frame.  After the standard external kinematic factors have been
removed, the contribution of one exchanged conformal family is obtained by
inserting the orthogonal projection to that family between the two states
created by the two operator pairs.  Denote these states by \(A\) and \(B\).
Let \(P_E\) be the spectral projection of the cylinder Hamiltonian
\(D_{\rm rad}\) onto the finite-dimensional eigenspace of radial energy \(E\)
inside the chosen family.  Rotations of \(S^{D-1}\) commute with
\(D_{\rm rad}\), so the level expansion is a matrix element of
\[
  r^{D_{\rm rad}}U(\alpha),
  \qquad
  \rho=r\ee^{\ii\alpha},
\]
where \(U(\alpha)\) is the unitary rotation relating the angular positions of
the two pairs.  For \(r\le q<1\), Cauchy--Schwarz gives
\[
  \sum_E r^E
  \left|
    \langle A,P_EU(\alpha)B\rangle
  \right|
  \le
  \left(\sum_E q^E\|P_EA\|^2\right)^{1/2}
  \left(\sum_E q^E\|P_EB\|^2\right)^{1/2}.
\]
The two sums on the right are
\[
  \|q^{D_{\rm rad}/2}A\|^2,\qquad
  \|q^{D_{\rm rad}/2}B\|^2 .
\]
By reflection positivity these norms are Euclidean correlation functions in
which the two operator pairs are separated by an annulus of radial modulus
\(q\).  The radial reconstruction hypothesis asserts exactly that such
separated Euclidean configurations define finite Hilbert-space vectors.  Hence
the displayed bound is finite and independent of the partial-sum cutoff and of
\(\alpha\).  This proves absolute convergence uniformly for \(r\le q\).

Finite multiplicity makes each energy-level contribution a finite matrix
element in the corresponding space of spherical harmonics; for scalar external
operators it is a finite linear combination of Gegenbauer polynomials in the
angular variable.  The same majorant therefore controls every compact
subdisk \(|\rho|\le q\), and the Weierstrass test gives local uniform
convergence of the radial series.  The identity with the conformal block
follows on the real radial domain from the complete-set construction; the
locally uniform limit is analytic in the radial variable, so the identity
theorem continues the equality to the \(\rho\)-disk and, through the inverse
map \(z=4\rho/(1+\rho)^2\), to the cut \(z\)-plane.
\end{proof}

At the crossing-symmetric Euclidean point \(z=\bar z=1/2\),
\begin{equation}
  \rho(1/2)
  =
  \frac{1-\sqrt{1/2}}{1+\sqrt{1/2}}
  =
  3-2\sqrt2\approx0.1716 .
  \label{eq:rho-crossing-symmetric-point}
\end{equation}
This small value, and for identical external scalars the additional even-level
selection discussed below, is the convergence mechanism behind efficient
radial-block evaluation near the bootstrap crossing point.

\begin{figure}[H]
  \centering
  \resizebox{0.88\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      insertion/.style={circle,fill=violet!80!black,inner sep=1.7pt},
      arrow/.style={-{Latex[length=2mm]},line width=0.8pt},
      guide/.style={blue!65!black,dashed,line width=0.75pt}
    }
    \begin{scope}
      \node[font=\small] at (0,2.05) {plane};
      \draw[arrow] (-2.0,0)--(2.0,0);
      \draw[arrow] (0,-1.45)--(0,1.55);
      \node[insertion] at (-1.35,0) {};
      \node[violet!80!black,below] at (-1.35,0) {\(-1\)};
      \node[insertion] at (1.35,0) {};
      \node[violet!80!black,below] at (1.35,0) {\(1\)};
      \draw[orange!85!black,line width=0.8pt] (0,0)--(0.70,0.55);
      \draw[orange!85!black,line width=0.8pt] (0,0)--(-0.70,-0.55);
      \node[insertion] at (0.70,0.55) {};
      \node[insertion] at (-0.70,-0.55) {};
      \node[violet!80!black,above right] at (0.70,0.55) {\(\rho\)};
      \node[violet!80!black,below left] at (-0.70,-0.55) {\(-\rho\)};
      \draw[guide] (0.58,0) arc[start angle=0,end angle=38,radius=0.58];
      \node[blue!65!black] at (0.78,0.20) {\(\alpha\)};
    \end{scope}

    \draw[arrow] (2.75,0)--(4.0,0);
    \node[align=center] at (3.38,0.55) {radial\\quantization};

    \begin{scope}[shift={(6.35,0)}]
      \node[font=\small] at (0,2.05) {cylinder};
      \draw[thick] (-1.45,-1.35)--(-1.45,1.35);
      \draw[thick] (1.45,-1.35)--(1.45,1.35);
      \draw[thick] (0,1.35) ellipse (1.45 and 0.35);
      \draw[thick] (0,-1.35) ellipse (1.45 and 0.35);
      \draw[guide] (0,0.70) ellipse (1.45 and 0.35);
      \draw[guide] (0,-0.25) ellipse (1.45 and 0.35);
      \node[insertion] at (1.45,0.70) {};
      \node[insertion] at (-1.45,0.70) {};
      \node[violet!80!black,right] at (1.45,0.70) {\((0,0)\)};
      \node[violet!80!black,left] at (-1.45,0.70) {\((0,\pi)\)};
      \node[insertion] at (0.65,-0.05) {};
      \node[insertion] at (-0.65,-0.45) {};
      \node[violet!80!black,right] at (0.65,-0.05)
        {\((\log r,\alpha)\)};
      \node[violet!80!black,below left] at (-0.65,-0.45)
        {\((\log r,\alpha+\pi)\)};
      \draw[green!45!black,<->,line width=0.8pt] (-1.12,-0.25)--(-1.12,0.70);
      \node[green!45!black,right] at (-1.12,0.20) {\(-\log r\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The radial variable \(\rho=r\ee^{\ii\alpha}\) puts the
  \(12\to34\) OPE channel into a cylinder frame.  Inserting a complete set of
  cylinder energy eigenstates gives a convergent expansion in powers of \(r\).}
  \label{fig:rho-cylinder-frame}
\end{figure}

Assume \(\phi\) is normalized by
\(\langle\!\langle\phi|\phi\rangle\!\rangle=1\).  In the cylinder frame,
inserting a complete basis of states gives
\begin{align}
  2^{-4\Delta_\phi}\mathcal G(u,v)
  &=
  \bra{\vac}
    \widetilde\phi(0,0)\widetilde\phi(0,\pi)
    \widetilde\phi(\log r,\alpha)
    \widetilde\phi(\log r,\alpha+\pi)
  \ket{\vac}
  \notag\\
  &=
  \sum_n r^{E_n}
  \bra{\vac}
    \widetilde\phi(0,0)\widetilde\phi(0,\pi)
  \ket{n}
  \bra{n}
    \widetilde\phi(0,\alpha)\widetilde\phi(0,\alpha+\pi)
  \ket{\vac}.
  \label{eq:cylinder-four-point-complete-set}
\end{align}
For a primary \(\Phi_i\) of spin \(s_i\) and dimension \(\Delta_i\), its
level-\(n\) descendants have energy \(\Delta_i+n\).  If
\(a=(\beta;N_1,\ldots,N_D)\), with \(\sum_\mu N_\mu=n\), labels a raw
descendant
\[
  \ket{\Phi_i,a}
  =
  \prod_{\mu=1}^D \widehat P_\mu^{\,N_\mu}
  \ket{\Phi_i,\beta},
\]
let
\[
  G^{\rm raw}_{ab}
  =
  \langle
    \Phi_i,a\,|\,\Phi_i,b
  \rangle
\]
be the level-\(n\) Gram form.  This raw Gram form is finite-dimensional at
fixed \(n\), but it can be singular.  Its null subspace is
\[
  \mathcal N_{i,n}
  =
  \{v\in W_{i,n}:\langle v|w\rangle=0\ \text{for every }w\in W_{i,n}\},
\]
where \(W_{i,n}\) is the raw level-\(n\) descendant space.  Reflection
positivity implies that every \(v\in\mathcal N_{i,n}\) has zero matrix element
against every physical state, by the Cauchy--Schwarz inequality.  Hence null
descendants are first quotiented:
\[
  Q_{i,n}=W_{i,n}/\mathcal N_{i,n}.
\]
Choose any basis \(\{\ket{e_A^{(n)}}\}_A\) of \(Q_{i,n}\), represented by raw
descendants, and define the nonsingular quotient Gram matrix
\[
  M^{(n)}_{AB}
  =
  \langle e_A^{(n)}|e_B^{(n)}\rangle .
\]
The level-\(n\) projector onto the exchanged conformal family is
\begin{equation}
  \Pi_{i,n}
  =
  \sum_{A,B}
  \ket{e_A^{(n)}}\,
  \bigl(M^{(n)}\bigr)^{-1}_{AB}\,
  \bra{e_B^{(n)}} .
  \label{eq:radial-block-level-projector}
\end{equation}
This expression is independent of the chosen quotient basis: under
\(\ket{e_A}\mapsto S_A{}^B\ket{e_B}\), the Gram matrix and its inverse
transform contragrediently.  The complete-set term at level \(n\) is therefore
\begin{equation}
  r^{\Delta_i+n}
  \sum_{A,B}
  \bra{\vac}
    \widetilde\phi(0,0)\widetilde\phi(0,\pi)
  \ket{e_A^{(n)}}\,
  \bigl(M^{(n)}\bigr)^{-1}_{AB}\,
  \bra{e_B^{(n)}}
    \widetilde\phi(0,\alpha)\widetilde\phi(0,\alpha+\pi)
  \ket{\vac}.
  \label{eq:radial-block-level-gram-inversion}
\end{equation}
The block is obtained by applying
\eqref{eq:radial-block-level-gram-inversion} to each finite level and then
summing over levels.  Thus no inverse of an infinite descendant Gram matrix is
being assumed; the inverse is taken after quotienting null states at each
energy level.  Conservation laws and other shortening conditions enter exactly
through the null subspaces \(\mathcal N_{i,n}\).  Consequently a block has a
radial expansion
\begin{equation}
  2^{-4\Delta_\phi}
  G_{\Delta_i,s_i}(u,v)
  =
  \sum_{n=0}^{\infty}
  A_{\Delta_i,s_i,n}(\alpha)\,r^{\Delta_i+n},
  \label{eq:block-radial-expansion}
\end{equation}
where each coefficient is fixed by conformal representation theory.

The leading coefficient is especially simple.  At level zero the only angular
dependence is the contraction of two symmetric traceless spin-\(s\) tensors
made from the two unit vectors separated by angle \(\alpha\).  Hence, with a
standard normalization of the block,
\begin{equation}
  G_{\Delta,s}(u,v)
  =
  r^\Delta
  C_s^{(D/2-1)}(\cos\alpha)
  +
  O(r^{\Delta+1}),
  \label{eq:block-leading-gegenbauer}
\end{equation}
where \(C_s^{(\lambda)}\) is the standard Gegenbauer polynomial.  Thus
\(C_s^{(1/2)}(\cos\alpha)=P_s(\cos\alpha)\) in \(D=3\), while
\[
  C_s^{(1)}(\cos\alpha)
  =
  \frac{\sin((s+1)\alpha)}{\sin\alpha}
\]
in \(D=4\).  The \(D=2\) angular functions are the separate limiting harmonic
\(\cos(s\alpha)\), since the Gegenbauer parameter \(\lambda=D/2-1\) vanishes.

\section{Reflection Positivity and OPE Coefficients}

Reflection positivity constrains the numerical OPE data.  Choose Hermitian
scalar primaries with two-point normalization
\[
  \langle \mathcal O_A(x)\mathcal O_B(0)\rangle
  =
  \frac{\delta_{AB}}{|x|^{2\Delta_A}} .
\]
With this normalization, phases of nondegenerate Hermitian primaries can be
chosen so that three-point coefficients of Hermitian scalar operators are
real.  If \(\phi\) is a Hermitian scalar primary, the contribution of a
primary family \(\mathcal O\) to the \(\phi\phi\phi\phi\) four-point function
has coefficient
\[
  \lambda_{\phi\phi\mathcal O}^2
  \ge0
\]
in the channel where the first pair of operators is radially inside the
second pair.  This nonnegative number is the squared norm of the projection of
the state \(\phi(x)\phi(0)\vac\) onto the conformal multiplet of
\(\mathcal O\).

When several operators with the same quantum numbers mix, the statement is
matrix-valued.  Let \(P_\Delta\) be the spectral projection onto a finite
degenerate subspace of fixed spin and scaling dimension.  For any coefficients
\(\alpha_i\),
\[
  \left\|
    P_\Delta
    \sum_i\alpha_i\,\mathcal O_i(x_i)\vac
  \right\|^2
  \ge0 .
\]
The corresponding OPE coefficients form a positive semidefinite matrix in
that degenerate subspace.  Positivity of OPE data is therefore a direct
Hilbert-space consequence of radial quantization, not a separate dynamical
assumption.

\section{Associativity}

Consider a Euclidean correlation function with three operators inside a sphere
that separates them from all other insertions.  If the insertion points admit
nested separating spheres, there are two ways to apply the OPE:
\[
  (\mathcal O_1\mathcal O_2)\mathcal O_3,
  \qquad
  \mathcal O_1(\mathcal O_2\mathcal O_3).
\]
Let \(S_{\rm out}\) be the outer separating sphere.  The product of the three
separated insertions defines a vector in the radial Hilbert space on
\(S_{\rm out}\), not a local operator.
The two nested OPE procedures are two ways of inserting complete spectral
resolutions of the same vector.  On a common nested radial domain the two
partial sums converge in Hilbert norm to that vector, hence have the same
matrix elements against exterior states.  Away from the common domain, the
statement becomes an analytic-continuation statement: each nested expansion
defines a holomorphic function in its own channel domain, and both are
restrictions of the same separated-point correlation function on the
connected component of complexified configuration space reached without
crossing coincident-point or light-cone singular hypersurfaces.

For identical scalar primaries \(\phi\) of dimension \(\Delta_\phi\), write
the four-point function in the form
\[
  \langle
    \phi(x_1)\phi(x_2)\phi(x_3)\phi(x_4)
  \rangle
  =
  \frac{\mathcal G(u,v)}
       {|x_{12}|^{2\Delta_\phi}|x_{34}|^{2\Delta_\phi}},
\]
where \(x_{ij}=x_i-x_j\) and the conformal cross-ratios are
\[
  u=\frac{x_{12}^2x_{34}^2}{x_{13}^2x_{24}^2},
  \qquad
  v=\frac{x_{14}^2x_{23}^2}{x_{13}^2x_{24}^2}.
\]
Introduce \(z,\bar z\) by \(u=z\bar z\) and
\(v=(1-z)(1-\bar z)\).  The \(12\to34\) channel is controlled by
\[
  \rho_s=\frac{z}{(1+\sqrt{1-z})^2},
  \qquad
  \bar\rho_s=\frac{\bar z}{(1+\sqrt{1-\bar z})^2},
\]
and its Euclidean radial block expansion converges on compact subsets of
\begin{equation}
  \mathcal D_s
  =
  \{(z,\bar z): |\rho_s|<1,\ |\bar\rho_s|<1\}.
  \label{eq:s-channel-radial-domain}
\end{equation}
The \(14\to23\) channel uses the crossed variables
\[
  \rho_t=\frac{1-z}{(1+\sqrt z)^2},
  \qquad
  \bar\rho_t=\frac{1-\bar z}{(1+\sqrt{\bar z})^2},
\]
and converges on compact subsets of
\begin{equation}
  \mathcal D_t
  =
  \{(z,\bar z): |\rho_t|<1,\ |\bar\rho_t|<1\}.
  \label{eq:t-channel-radial-domain}
\end{equation}
These domains are channel domains, not identical power-series neighborhoods:
\(\mathcal D_s\) contains the \(u\to0\) OPE boundary, while
\(\mathcal D_t\) contains the \(v\to0\) OPE boundary.  Their Euclidean
intersection is nonempty; for example \(0<z=\bar z<1\) lies in both domains.
On that intersection the two convergent Hilbert-space expansions compute the
same four-point function.  The identity theorem then identifies their
holomorphic continuations on every connected component where both
continuations are defined.

The crossing equation itself follows from the permutation symmetry of the
identical Euclidean scalar correlator.  Exchanging \(x_2\) and \(x_4\) sends
\((u,v)\) to \((v,u)\).  The same separated-point function may therefore be
written as
\[
  \frac{\mathcal G(u,v)}
       {|x_{12}|^{2\Delta_\phi}|x_{34}|^{2\Delta_\phi}}
  =
  \frac{\mathcal G(v,u)}
       {|x_{14}|^{2\Delta_\phi}|x_{23}|^{2\Delta_\phi}} .
\]
Since
\[
  \frac{u}{v}
  =
  \frac{x_{12}^2x_{34}^2}{x_{14}^2x_{23}^2},
\]
this equality is equivalent to
\begin{equation}
  v^{\Delta_\phi}\mathcal G(u,v)
  =
  u^{\Delta_\phi}\mathcal G(v,u).
  \label{eq:identical-scalar-crossing}
\end{equation}
If \(\phi\) is Hermitian and the exchanged primaries are orthonormal, the
\(12\to34\) expansion takes the form
\begin{equation}
  \mathcal G(u,v)
  =
  1+
  \sum_{\mathcal O\ne\mathbf 1}
  \lambda_{\phi\phi\mathcal O}^2
  G_{\Delta_{\mathcal O},s_{\mathcal O}}(u,v),
  \qquad
  \lambda_{\phi\phi\mathcal O}^2\ge0.
  \label{eq:identical-scalar-positive-block-expansion}
\end{equation}
The identity block gives the constant \(1\).  The \(14\to23\) expansion gives
the same formula with \((u,v)\mapsto(v,u)\), valid in the channel domain
\(\mathcal D_t\).  Equation~\eqref{eq:identical-scalar-crossing} is therefore
the equality of two OPE-defined holomorphic functions after both have been
identified with the analytic continuation of one Euclidean separated-point
correlator.  It is not obtained by comparing two unrelated power series at
their distinct expansion corners.

\section{Worked Example: Generalized-Free Scalar Crossing}
\label{sec:gff-four-point-crossing}

The preceding discussion becomes completely concrete for a scalar
generalized-free datum.  Assume \(D>2\) and
\(\Delta_\phi\ge (D-2)/2\), or \(D=2\) and \(\Delta_\phi\ge0\) for a spin-zero
global conformal scalar.  Normalize
\[
  \langle \phi(x)\phi(y)\rangle_{\rm GFF}
  =
  |x-y|^{-2\Delta_\phi}.
\]
The separated-point \(2m\)-point functions are defined by Wick pairings,
\begin{equation}
  \left\langle
    \phi(x_1)\cdots\phi(x_{2m})
  \right\rangle_{\rm GFF}
  =
  \sum_{\pi\in{\rm Pair}(2m)}
  \prod_{\{i,j\}\in\pi}|x_i-x_j|^{-2\Delta_\phi},
  \qquad
  \langle \phi(x_1)\cdots\phi(x_{2m+1})\rangle_{\rm GFF}=0 .
  \label{eq:gff-wick-definition}
\end{equation}
Here \({\rm Pair}(2m)\) denotes the set of partitions of
\(\{1,\ldots,2m\}\) into unordered two-element subsets.  In the special case
\(D>2\) and \(\Delta_\phi=(D-2)/2\), the quotient by the null descendant
\(\partial^2\phi\) gives the ordinary massless free scalar CFT.  For larger
\(\Delta_\phi\) the same formula defines a generalized-free scalar sector.
The statements below use the scalar module generated by \(\phi\), the Wick
pairing rule, and radial reflection positivity.

In radial quantization, \(\phi(0)\ket{\vac}\) generates the scalar
lowest-weight conformal module \(V_{\Delta_\phi}\), quotienting null vectors
when the unitarity bound is saturated.  Equation~\eqref{eq:gff-wick-definition}
makes the \(m\)-particle Gram matrix the permanent Gram matrix on
\(\operatorname{Sym}^m V_{\Delta_\phi}\), meaning that its entries are sums
over pairings with no sign character.  Thus the reconstructed radial
Hilbert space of this sector is the symmetric Fock space over
\(V_{\Delta_\phi}\), before Hilbert completion and after the same null-vector
quotient.  This finite Gram-matrix statement is the reason the OPE below has
nonnegative squared coefficients.

For four points Wick factorization gives the three pairings
\begin{equation}
  \begin{split}
  \left\langle
    \phi(x_1)\phi(x_2)\phi(x_3)\phi(x_4)
  \right\rangle_{\rm GFF}
  &=
  \frac{1}{|x_{12}|^{2\Delta_\phi}|x_{34}|^{2\Delta_\phi}}
  +\frac{1}{|x_{13}|^{2\Delta_\phi}|x_{24}|^{2\Delta_\phi}}
  \\
  &\hspace{1.5em}
  +\frac{1}{|x_{14}|^{2\Delta_\phi}|x_{23}|^{2\Delta_\phi}} .
  \end{split}
  \label{eq:gff-four-point-wick}
\end{equation}
With the four-point prefactor and cross-ratios used in
\eqref{eq:identical-scalar-crossing}, the reduced correlator is therefore
\begin{equation}
  \mathcal G_{\rm GFF}(u,v)
  =
  1+u^{\Delta_\phi}
  +\left(\frac{u}{v}\right)^{\Delta_\phi}.
  \label{eq:gff-reduced-four-point}
\end{equation}
The crossing relation is now an algebraic identity:
\begin{equation}
  v^{\Delta_\phi}\mathcal G_{\rm GFF}(u,v)
  =
  v^{\Delta_\phi}
  +(uv)^{\Delta_\phi}
  +u^{\Delta_\phi}
  =
  u^{\Delta_\phi}
  +(uv)^{\Delta_\phi}
  +v^{\Delta_\phi}
  =
  u^{\Delta_\phi}\mathcal G_{\rm GFF}(v,u).
  \label{eq:gff-crossing-check}
\end{equation}

The \(12\to34\) OPE contains the identity and bilinear conformal primaries.
Let \(z^\mu\) be a null auxiliary vector, \(z^2=0\), used to encode symmetric
traceless spin.  For \(n\in\mathbb Z_{\ge0}\) and
\(\ell\in2\mathbb Z_{\ge0}\), define
\begin{equation}
  \mathcal O_{n,\ell}(x,z)
  =
  \mathcal N_{n,\ell}\,
  \Pi_{\rm prim}
  \left[
    :\phi(x_1)
    \bigl(z\cdot(\partial_{x_1}-\partial_{x_2})\bigr)^\ell
    \bigl((\partial_{x_1}-\partial_{x_2})^2\bigr)^n
    \phi(x_2):
  \right]_{x_1=x_2=x}.
  \label{eq:gff-double-trace-primary-definition}
\end{equation}
The colons denote Wick subtraction with respect to
\eqref{eq:gff-wick-definition}.  The projection \(\Pi_{\rm prim}\) is
defined as follows.  At fixed total relative derivative degree
\(N=2n+\ell\), the rotational spin-\(\ell\) bilinears form a finite-dimensional
inner-product space.  Total derivatives are \(P_\mu\)-descendants of lower
energy states.  Since radial reflection makes \(K_\mu=P_\mu^\dagger\), the
primary subspace is the orthogonal complement of the descendant subspace,
equivalently the kernel of every \(K_\mu\).  The finite orthogonal projection
onto this kernel is \(\Pi_{\rm prim}\).  Homogeneous polynomials in the
relative derivative decompose as
\[
  \operatorname{Sym}^N(\R^D)
  =
  \bigoplus_{j=0}^{\lfloor N/2\rfloor}
  (q^2)^j\,\mathcal H_{N-2j},
\]
where \(\mathcal H_m\) is the space of degree-\(m\) harmonic polynomials.
For \(N=2n+\ell\), the quotient by total derivatives contains one copy of the
spin-\(\ell\) harmonic representation.  Its projection has nonzero Wick norm,
so it gives one
orthonormal primary family for each \((n,\ell)\) that survives the null
quotient.  Exchange of the two identical fields sends the relative derivative
to its negative, so only even \(\ell\) survives in the symmetric product.

The normalization \(\mathcal N_{n,\ell}\) is chosen so that
\[
  \langle
    \mathcal O_{n,\ell}(x,z)
    \mathcal O_{n',\ell'}(0,z')
  \rangle
  =
  0
  \quad\hbox{unless}\quad
  (n,\ell)=(n',\ell'),
\]
and the nonzero two-point function has unit coefficient in the spin-\(\ell\)
conformal normalization used for the blocks in this chapter.  The
corresponding scaling dimension and twist are
\begin{equation}
  \Delta_{n,\ell}=2\Delta_\phi+2n+\ell,
  \qquad
  \tau_{n,\ell}=\Delta_{n,\ell}-\ell=2\Delta_\phi+2n .
  \label{eq:gff-double-trace-dimensions}
\end{equation}
The dimension formula follows directly from radial energy: two scalar
one-particle states contribute \(2\Delta_\phi\), each derivative contributes
one unit, and the primary projection only subtracts states with the same
total energy.

With orthonormal primaries the \(12\to34\) channel expansion is
\begin{equation}
  \mathcal G_{\rm GFF}(u,v)
  =
  1+
  \sum_{n=0}^{\infty}
  \sum_{\ell\in2\mathbb Z_{\ge0}}
  a^{\rm GFF}_{n,\ell}\,
  G_{2\Delta_\phi+2n+\ell,\ell}(u,v),
  \qquad
  a^{\rm GFF}_{n,\ell}
  =
  \lambda_{\phi\phi\mathcal O_{n,\ell}}^2\ge0 ,
  \label{eq:gff-s-channel-block-expansion}
\end{equation}
valid in the \(s\)-channel radial domain
\(\mathcal D_s\) of \eqref{eq:s-channel-radial-domain}.  This formula is the
radial spectral resolution of the vector
\(\phi(x_1)\phi(x_2)\ket{\vac}\).  Its identity component is the contraction
\(|x_{12}|^{-2\Delta_\phi}\ket{\vac}\); its normal-ordered component lies in
\(\operatorname{Sym}^2V_{\Delta_\phi}\), whose primary decomposition was
constructed in the preceding paragraph.  Pairing this convergent vector
expansion with the exterior state
\(\phi(x_3)\phi(x_4)\ket{\vac}\) gives
\eqref{eq:gff-s-channel-block-expansion}.  Since the same matrix element was
already computed by the finite Wick sum \eqref{eq:gff-four-point-wick},
uniqueness of the orthogonal spectral resolution identifies the block sum
with \eqref{eq:gff-reduced-four-point} throughout \(\mathcal D_s\).

Equivalently, with spin indices encoded by the auxiliary null vector or
displayed explicitly at level zero, the OPE has the radial Hilbert-space form
\begin{equation}
  \phi(x)\phi(0)\ket{\vac}
  =
  |x|^{-2\Delta_\phi}\ket{\vac}
  +
  \sum_{n=0}^{\infty}
  \sum_{\ell\in2\mathbb Z_{\ge0}}
  \lambda_{\phi\phi\mathcal O_{n,\ell}}\,
  \sum_{r=0}^{\infty}
  \mathcal C^{(r)}_{n,\ell}(x,P)\,
  \mathcal O_{n,\ell}(0)\ket{\vac},
  \label{eq:gff-primary-ope-vector}
\end{equation}
where \(P_\mu\) acts on the primary state to generate descendants.  The
level-\(r\) coefficient is a homogeneous polynomial-valued differential
operator
\[
  \mathcal C^{(r)}_{n,\ell}(x,P)
  =
  \sum_{|\alpha|=r}
  c^{(n,\ell)}_{\alpha}(x)\,P^\alpha,
\]
obtained by applying the finite projector
\eqref{eq:radial-block-level-projector} at descendant level \(r\).  Its
primary-level term is
\[
  \mathcal C^{(0)}_{n,\ell}(x,P)\,
  \mathcal O_{n,\ell}(0)
  =
  |x|^{2n}
  x_{\mu_1}\cdots x_{\mu_\ell}
  \mathcal O_{n,\ell}^{\mu_1\cdots\mu_\ell}(0).
\]
The power of \(|x|\) is fixed by dimension:
\(\Delta_{n,\ell}-2\Delta_\phi-\ell=2n\).

The first nontrivial scalar primary illustrates the normalization.  Put
\[
  \mathcal O_{0,0}(x)=\frac{1}{\sqrt2}:\phi(x)^2: .
\]
Wick contraction gives
\[
  \langle :\phi(x)^2::\phi(0)^2:\rangle
  =
  \frac{2}{|x|^{4\Delta_\phi}},
  \qquad
  \langle \phi(x_1)\phi(x_2):\phi(x_3)^2:\rangle
  =
  \frac{2}{|x_{13}|^{2\Delta_\phi}|x_{23}|^{2\Delta_\phi}} .
\]
Hence
\begin{equation}
  \lambda_{\phi\phi\mathcal O_{0,0}}=\sqrt2,
  \qquad
  a^{\rm GFF}_{0,0}=2 .
  \label{eq:gff-leading-scalar-coefficient}
\end{equation}
The first two terms of \eqref{eq:gff-primary-ope-vector} are therefore
\[
  \phi(x)\phi(0)\ket{\vac}
  =
  |x|^{-2\Delta_\phi}\ket{\vac}
  +\sqrt2\,\mathcal O_{0,0}(0)\ket{\vac}
  +\cdots ,
\]
where the omitted terms are precisely the positive-level terms in
\eqref{eq:gff-primary-ope-vector}.
The remaining terms in the two functions
\(u^{\Delta_\phi}\) and \((u/v)^{\Delta_\phi}\), expanded near the
\(s\)-channel OPE boundary, are reproduced by the descendants of
\(\mathcal O_{0,0}\) together with the higher-spin and higher-\(n\) bilinear
primaries in \eqref{eq:gff-s-channel-block-expansion}.

The \(14\to23\) OPE is the same representation-theoretic decomposition in the
crossed channel.  Written with the \(12\to34\) prefactor, it reads
\begin{equation}
  \mathcal G_{\rm GFF}(u,v)
  =
  \left(\frac{u}{v}\right)^{\Delta_\phi}
  \left[
    1+
    \sum_{n=0}^{\infty}
    \sum_{\ell\in2\mathbb Z_{\ge0}}
    a^{\rm GFF}_{n,\ell}\,
    G_{2\Delta_\phi+2n+\ell,\ell}(v,u)
  \right],
  \label{eq:gff-t-channel-block-expansion}
\end{equation}
valid in \(\mathcal D_t\).  The identity operator in the crossed channel is
responsible for the term \((u/v)^{\Delta_\phi}\) in
\eqref{eq:gff-reduced-four-point}; the crossed bilinear tower sums to
\((u/v)^{\Delta_\phi}(v^{\Delta_\phi}+(v/u)^{\Delta_\phi})
=u^{\Delta_\phi}+1\).  Thus the two OPE channels are two convergent radial
expansions of the same Wick-pairing function, and
\eqref{eq:gff-crossing-check} is the prefactor form of that equality.

\section{A First Consequence of Crossing and Positivity}

The simplest use of \eqref{eq:identical-scalar-crossing} is to see why a
unitary CFT cannot have an arbitrarily large gap above the identity in the
\(\phi\times\phi\) OPE while keeping \(\Delta_\phi\) fixed.  This subsection
is deliberately a controlled approximation, not a sharp bootstrap theorem:
we keep the leading radial behavior of heavy exchanged blocks near the
crossing-symmetric point and state the assumptions used.

Take \(z=\bar z\in(0,1)\), so that \(u=z^2\) and \(v=(1-z)^2\).  The crossing
equation becomes
\begin{equation}
  (1-z)^{2\Delta_\phi}\mathcal G(z)
  =
  z^{2\Delta_\phi}\mathcal G(1-z),
  \label{eq:real-line-identical-crossing}
\end{equation}
where \(\mathcal G(z)\) abbreviates \(\mathcal G(z^2,(1-z)^2)\).  Using
\eqref{eq:identical-scalar-positive-block-expansion},
\begin{equation}
  \bigl[(1-z)^{2\Delta_\phi}-z^{2\Delta_\phi}\bigr]
  +
  \sum_{\mathcal O\ne\mathbf1}
  \lambda_{\mathcal O}^2
  \left[
    (1-z)^{2\Delta_\phi}G_{\mathcal O}(z)
    -
    z^{2\Delta_\phi}G_{\mathcal O}(1-z)
  \right]
  =
  0.
  \label{eq:crossing-expanded-identity-plus-blocks}
\end{equation}
For \(0<z<1\), define
\[
  r(z)=\frac{z}{(1+\sqrt{1-z})^2}.
\]
If the exchanged dimensions contributing above the identity are large, and if
we keep only the leading radial behavior
\[
  G_{\mathcal O}(z)
  \simeq
  r(z)^{\Delta_{\mathcal O}}
\]
on this line, then \eqref{eq:crossing-expanded-identity-plus-blocks} is
approximated by
\begin{equation}
  \bigl[(1-z)^{2\Delta_\phi}-z^{2\Delta_\phi}\bigr]
  +
  \sum_{\mathcal O\ne\mathbf1}
  \lambda_{\mathcal O}^2
  \left[
    (1-z)^{2\Delta_\phi}r(z)^{\Delta_{\mathcal O}}
    -
    z^{2\Delta_\phi}r(1-z)^{\Delta_{\mathcal O}}
  \right]
  =
  0.
  \label{eq:large-gap-approx-crossing}
\end{equation}
The crossing-symmetric point is \(z=1/2\), where
\[
  r(1/2)=3-2\sqrt2.
\]
Set \(z=1/2+x\) and expand \eqref{eq:large-gap-approx-crossing} in \(x\).
The identity contribution is
\[
  (1-z)^{2\Delta_\phi}-z^{2\Delta_\phi}
  =
  -\Delta_\phi 2^{3-2\Delta_\phi}
  \left[
    x+\frac43(\Delta_\phi-1)(2\Delta_\phi-1)x^3+\cdots
  \right].
\]
The coefficient follows directly from
\((1/2-x)^{2\Delta_\phi}-(1/2+x)^{2\Delta_\phi}
=2^{-2\Delta_\phi}[(1-2x)^{2\Delta_\phi}-(1+2x)^{2\Delta_\phi}]\);
the odd cubic term is
\(-2^{4-2\Delta_\phi}\binom{2\Delta_\phi}{3}x^3\), which becomes the
displayed \(\frac43(\Delta_\phi-1)(2\Delta_\phi-1)\) after the common factor
\(-\Delta_\phi2^{3-2\Delta_\phi}\) is extracted.
For a heavy exchanged primary, the leading odd part of the approximate block
contribution has the form
\[
  \widetilde\lambda_{\mathcal O}^{\,2}
  \left[
    x+\frac43\Delta_{\mathcal O}^2x^3+\cdots
  \right],
  \qquad
  \widetilde\lambda_{\mathcal O}^{\,2}\ge0,
\]
where the positive factor relating
\(\widetilde\lambda_{\mathcal O}^{\,2}\) to
\(\lambda_{\mathcal O}^2\) includes
\((3-2\sqrt2)^{\Delta_{\mathcal O}}\) and other positive constants.  Matching
the coefficients of \(x\) and \(x^3\) gives, in this approximation,
\[
  \sum_{\mathcal O\ne\mathbf1}
  \widetilde\lambda_{\mathcal O}^{\,2}
  \left[
    \Delta_{\mathcal O}^2
    -
    (\Delta_\phi-1)(2\Delta_\phi-1)
  \right]
  =
  0.
\]
Since every coefficient in the sum is nonnegative, not all exchanged
dimensions can be much larger than
\[
  \sqrt{(\Delta_\phi-1)(2\Delta_\phi-1)}.
\]
The robust conclusion is qualitative: crossing plus reflection positivity
forces nontrivial operator content in the \(\phi\times\phi\) OPE at dimensions
controlled by \(\Delta_\phi\).  A rigorous numerical or analytic bound
requires the full conformal blocks and lies beyond this first approximation.
