\chapter{Stress Tensor, Weyl Structure, and Improvement}
\label{ch:volume-iii-stress-tensor-weyl}
\ConventionsEuclideanCFT


\section{The Stress Tensor from Background Metrics}

\begin{definition}[Metric-source stress tensor]
\label{def:metric-source-stress-tensor}
Let a Euclidean local theory be coupled to a smooth background metric
\(g_{\mu\nu}\) on a \(D\)-dimensional manifold \(M\).  The stress tensor is
defined by the first variation of the action or of the renormalized generating
functional with respect to the metric.  This chapter uses the
contravariant-metric convention
\begin{equation}
  \delta S[\Phi;g]
  =
  \frac12
  \int_M \dd^D x\,\sqrt g\,
  T_{\mu\nu}(x)\,\delta g^{\mu\nu}(x),
  \label{eq:stress-tensor-metric-variation}
\end{equation}
where \(\Phi\) denotes the dynamical fields and
\(g=\det(g_{\mu\nu})\).  Since
\(\delta g^{\mu\nu}=-g^{\mu\rho}g^{\nu\sigma}\delta g_{\rho\sigma}\), the same
convention is equivalently
\[
  \delta S[\Phi;g]
  =
  -\frac12
  \int_M \dd^D x\,\sqrt g\,
  T^{\mu\nu}(x)\,\delta g_{\mu\nu}(x).
\]
Equivalently, at the level of functional derivatives with all other sources
held fixed,
\begin{equation}
  T^{\mu\nu}(x)
  =
  -\frac{2}{\sqrt g(x)}
  \frac{\delta}{\delta g_{\mu\nu}(x)}
  =
  \frac{2}{\sqrt g(x)}
  g^{\mu\rho}(x)g^{\nu\sigma}(x)
  \frac{\delta}{\delta g^{\rho\sigma}(x)} .
  \label{eq:stress-tensor-covariant-contravariant-derivatives}
\end{equation}
The first differential operator acts on the functional written as a function of
\(g_{\mu\nu}\), and the second acts on the same functional written as a function
of \(g^{\mu\nu}\).  This identity is only the chain rule applied to
\(\delta g^{\mu\nu}=-g^{\mu\rho}g^{\nu\sigma}\delta g_{\rho\sigma}\).
This is the convention used by the metric derivative of the Euclidean
source functional.  At a regulator scale for which the metric-dependent
functional integral is a well-defined regulated integral, let
\[
  Z[g]=\int \mathcal D_g\Phi\,\exp\{-S[\Phi;g]\},
  \qquad
  \mathcal W[g]=-\log Z[g].
\]
For variations supported away from boundary components, or with boundary
conditions fixed so that no boundary term is produced, differentiating the
regulated integral gives
\[
  \delta\mathcal W[g]
  =
  \langle\delta S[\Phi;g]\rangle_g
  =
  -\frac12
  \int_M\dd^D x\,\sqrt g\,
  \langle T^{\mu\nu}(x)\rangle_g\,\delta g_{\mu\nu}(x).
\]
Therefore
\[
  \langle T^{\mu\nu}(x)\rangle_g
  =
  -\frac{2}{\sqrt g(x)}
  \frac{\delta\mathcal W[g]}{\delta g_{\mu\nu}(x)}.
\]
After subtracting local metric counterterms, the same identity defines the
renormalized source chart used in
\eqref{eq:cft-source-derivative-conventions}.  If instead one writes
\(K[g]=\log Z[g]\), the right-hand side carries the opposite sign; the
difference is only the choice between the logarithm of the partition function
and the Euclidean effective action.
With this sign convention, evaluating \(T_{\mu\nu}\) on the flat metric gives
the translation Noether tensor after adding the divergence term that makes the
Noether tensor symmetric.
\end{definition}

\begin{remark}[Curved extensions and flat stress tensors]
\label{rem:curved-extensions-flat-stress}
A flat classical action can have several curved extensions.  For example,
\[
  S[\phi]
  =
  \int_{\R^D}\dd^Dx\,
  \left(
  \frac12\partial_\mu\phi\,\partial^\mu\phi
  +
  \frac12m^2\phi^2
  \right)
\]
may be extended as
\[
  S[\phi;g]
  =
  \int_M\dd^Dx\,\sqrt g\,
  \left(
  \frac12g^{\mu\nu}\partial_\mu\phi\,\partial_\nu\phi
  +
  \frac12m^2\phi^2
  +
  \text{local curvature couplings}
  \right).
\]
Curvature couplings that vanish on flat space still affect \(T_{\mu\nu}\),
because \(T_{\mu\nu}\) is defined by varying \(g_{\mu\nu}\) before setting
\(g_{\mu\nu}=\delta_{\mu\nu}\).
\end{remark}

\paragraph{Diffeomorphism Ward identity for metric stress tensors.}
If the theory is invariant under diffeomorphisms of the background, then an
infinitesimal vector field \(\xi^\mu\) gives
\[
  \delta_\xi g^{\mu\nu}
  =
  -\nabla^\mu\xi^\nu-\nabla^\nu\xi^\mu .
\]
Substituting this variation into
\eqref{eq:stress-tensor-metric-variation} and integrating by parts gives
\begin{equation}
  \nabla_\mu T^{\mu\nu}=0
  \label{eq:stress-tensor-covariant-conservation}
\end{equation}
as an operator equation away from insertions.  In flat space this becomes
\(\partial_\mu T^{\mu\nu}=0\).  Symmetry of \(T_{\mu\nu}\) follows from
variation with respect to the symmetric tensor \(g_{\mu\nu}\).
The calculation is the source-form of Noether's identity.  Insert
\(\delta_\xi g^{\mu\nu}=-\nabla^\mu\xi^\nu-\nabla^\nu\xi^\mu\) into
Definition~\ref{def:metric-source-stress-tensor}.  The symmetry of
\(T_{\mu\nu}\) gives
\[
  \delta_\xi S
  =
  -\int_M\dd^Dx\,\sqrt g\,T_{\mu\nu}\nabla^\mu\xi^\nu .
\]
Integrating by parts gives
\[
  \delta_\xi S
  =
  \int_M\dd^Dx\,\sqrt g\,(\nabla_\mu T^{\mu\nu})\xi_\nu
\]
up to the boundary term excluded by the support or boundary conditions.  Since
the background-diffeomorphism variation is zero and \(\xi\) is arbitrary, the
local identity \(\nabla_\mu T^{\mu\nu}=0\) follows away from operator
insertions.  At coincident insertions the same source argument gives the
familiar contact terms; their precise form depends on which inserted
operators transform under the diffeomorphism and on the chosen contact-term
extension of the correlator.

\section{Metric Perturbations and Stress Insertions}

Set \(g_{\mu\nu}=\delta_{\mu\nu}+\delta g_{\mu\nu}\), with \(\delta g\) smooth
and compactly supported.  To first order, the Euclidean path integral changes
by the insertion
\begin{equation}
  \exp[-\delta_g S]
  =
  \exp\left[
  -\frac12\int_{\R^D}\dd^Dx\,
  \delta g^{\mu\nu}(x)T_{\mu\nu}(x)
  \right],
  \label{eq:stress-insertion-small-metric}
\end{equation}
where indices on \(\delta g^{\mu\nu}\) are raised with the flat metric at this
order.  Thus, for an operator product \(X\) supported away from
\(\operatorname{supp}(\delta g)\),
\[
  \langle X\rangle_{\delta+\delta g}
  =
  \langle X\rangle_{\delta}
  -
  \frac12\int\dd^Dx\,\delta g^{\mu\nu}(x)
  \langle T_{\mu\nu}(x)X\rangle_{\delta}
  +
  O(\delta g^2).
\]
Higher metric variations require a prescription for coincident products of
stress tensors.  The short-distance singularities in
\(T_{\mu\nu}(x)T_{\rho\sigma}(y)\) and the corresponding contact terms are part
of the renormalized definition of the theory on curved backgrounds.  This is a
local ultraviolet statement: in a massless theory the distributional extension
of stress-tensor products is used only after the infrared state or regulator
has made the separated-point correlators meaningful on the test functions
under consideration.  In even dimensions these contact terms include the Weyl
anomaly data.
Expand the metric-source action to first order in \(\delta g\) using
Definition~\ref{def:metric-source-stress-tensor}.  Since
\(\delta g^{\mu\nu}=-\delta g_{\mu\nu}+O(\delta g^2)\) around the flat metric,
the first-order change of the Boltzmann weight is the exponential insertion in
\eqref{eq:stress-insertion-small-metric}.  Expanding that exponential to first
order gives the displayed variation of \(\langle X\rangle\).  Higher
functional derivatives differentiate products of metric variations at
coincident points, so their definition requires the local extension of
stress-tensor products as distributions.

\section{Weyl Transformations}

\begin{definition}[Weyl transformation and trace insertion]
\label{def:weyl-transformation-trace-insertion}
A Weyl transformation is a local rescaling of the background metric:
\[
  g_{\mu\nu}(x)\mapsto \ee^{2\omega(x)}g_{\mu\nu}(x),
\]
where \(\omega\) is a smooth real function.  Infinitesimally,
\[
  \delta_\omega g_{\mu\nu}=2\omega g_{\mu\nu},
  \qquad
  \delta_\omega g^{\mu\nu}=-2\omega g^{\mu\nu}.
\]
The corresponding metric variation gives
\[
  \delta_\omega S
  =
  -
  \int_M\dd^D x\,\sqrt g\,
  \omega(x)\,T^\mu{}_\mu(x).
\]
\end{definition}
\begin{theorem}[Flat-space trace condition from Weyl invariance]
\label{thm:flat-space-weyl-trace-condition}
Assume that the renormalized theory is coupled to a background metric, that
the stress tensor is defined by metric variation as in
\eqref{eq:stress-insertion-small-metric}, and that a choice of field Weyl
transformations and allowed improvements has been made for which the
renormalized functional is invariant under compactly supported Weyl
rescalings in flat space, apart from curvature anomaly terms and contact terms
on coincident insertions.  Then, at separated flat-space points,
\begin{equation}
  T^\mu{}_\mu=0
  \label{eq:weyl-trace-zero}
\end{equation}
\end{theorem}

\begin{proof}
The action may assign Weyl transformations to the dynamical fields \(\Phi\).
After including these field variations and the chosen improvement terms, Weyl
invariance says that the first variation of every separated correlator under
\(\delta_\omega g_{\mu\nu}=2\omega g_{\mu\nu}\) vanishes, modulo the stated
anomaly and contact contributions.  The metric-variation formula above gives
\[
  \delta_\omega\langle X\rangle
  =
  -\int\dd^Dx\,\omega(x)
  \left\langle T^\mu{}_\mu(x)X\right\rangle
\]
for every operator product \(X\) supported away from
\(\operatorname{supp}(\omega)\).  Since \(\omega\) is an arbitrary smooth test
function in the separated region, the distributional insertion of
\(T^\mu{}_\mu\) vanishes there.  This is the local operator equation
\eqref{eq:weyl-trace-zero} at separated flat-space points.
\end{proof}

The background metric is an external source for stress-tensor insertions and
Ward identities.

In even dimensions, the renormalized generating functional may transform
under a Weyl rescaling by a local curvature functional.  At separated
flat-space points the conformal Ward identities are governed by
\eqref{eq:weyl-trace-zero}; the anomaly contributes contact terms and
curved-space response data.  The present volume records this contact-term
distinction; the special two-dimensional extension of conformal symmetry is
deferred to a separate treatment.

\section{Weyl Anomaly Density and Contact Ward Identities}

\begin{definition}[Weyl anomaly density]
\label{def:weyl-anomaly-density}
Define the trace anomaly density by
\begin{equation}
  \delta_\omega\mathcal W[g]
  =
  -\int_M\dd^D x\,\sqrt g\,\omega(x)\,\mathcal A[g](x),
  \qquad
  \mathcal A[g](x)=\left\langle T^\mu{}_\mu(x)\right\rangle_g .
  \label{eq:weyl-anomaly-density-definition}
\end{equation}
This sign follows from the convention
\(\langle T^{\mu\nu}\rangle_g=-2g^{-1/2}\delta\mathcal W/\delta g_{\mu\nu}\)
and \(\delta_\omega g_{\mu\nu}=2\omega g_{\mu\nu}\).  The anomaly density is a
local scalar of Weyl weight \(D\), defined modulo the Weyl variation of local
metric counterterms.  Its functional derivatives are contact terms in
stress-tensor Ward identities.  More explicitly, let
\[
  X_{n-1}
  =
  T^{\rho_1\sigma_1}(x_1)\cdots
  T^{\rho_{n-1}\sigma_{n-1}}(x_{n-1}).
\]
Tracing one stress tensor in an \(n\)-point distribution gives, away from
separated-point terms,
\[
  \begin{aligned}
  g_{\mu\nu}(x)
  \left\langle T^{\mu\nu}(x)X_{n-1}\right\rangle
  =
  \left.
  \frac{\delta^{\,n-1}\mathcal A[g](x)}
       {\prod_{j=1}^{n-1}\delta g_{\rho_j\sigma_j}(x_j)}
  \right|_{g=\delta}
  +\text{index-variation contact terms}.
  \end{aligned}
\]
Here the denominator denotes the iterated functional derivative with respect
to the displayed metric components.  Thus anomaly coefficients are part of
the metric-source definition of the stress-tensor correlators, even though
\(T^\mu{}_\mu=0\) at separated flat-space points.
\end{definition}

\section{Classification of Conformal Anomalies}

\paragraph{Wess--Zumino consistency for Weyl anomalies.}
The local anomaly density is constrained because Weyl transformations form an
abelian group.  Let
\[
  \mathbf A_\omega[g]
  :=
  \int_M\dd^D x\,\sqrt g\,\omega(x)\mathcal A[g](x).
\]
The identity
\([\delta_{\omega_1},\delta_{\omega_2}]\mathcal W[g]=0\) gives the
Wess--Zumino consistency condition
\begin{equation}
  \delta_{\omega_1}\mathbf A_{\omega_2}[g]
  -
  \delta_{\omega_2}\mathbf A_{\omega_1}[g]
  =
  0.
  \label{eq:weyl-wess-zumino-consistency}
\end{equation}
Changing the renormalized source functional by a local metric counterterm
\(\mathcal W[g]\mapsto\mathcal W[g]+B[g]\) changes
\(\mathbf A_\omega[g]\) by \(-\delta_\omega B[g]\).  Therefore the intrinsic
trace anomaly is the cohomology of the Weyl differential on local metric
functionals: solutions of
\eqref{eq:weyl-wess-zumino-consistency} modulo Weyl variations of local
counterterms.
This is the source-functional form of the abelian Weyl algebra.  Apply two
Weyl variations to the same renormalized functional \(\mathcal W[g]\).
Since local Weyl rescalings commute,
\([\delta_{\omega_1},\delta_{\omega_2}]\mathcal W[g]=0\).  Using
Definition~\ref{def:weyl-anomaly-density}, this commutator is precisely
\(\delta_{\omega_1}\mathbf A_{\omega_2}
-\delta_{\omega_2}\mathbf A_{\omega_1}\), giving
\eqref{eq:weyl-wess-zumino-consistency}.  If
\(\mathcal W\) is shifted by a local counterterm \(B[g]\), its anomaly shifts
by the Weyl variation of that counterterm.  Therefore only the cohomology class
of \(\mathbf A_\omega\) is intrinsic.

\begin{quotedtheorem}[Parity-even local Weyl anomaly classes]
\label{qthm:parity-even-local-weyl-anomaly-classes}
In the parity-even metric sector on a closed manifold without boundary, the
bulk solutions in even dimension \(D=2n\) have the form
\begin{equation}
  \mathcal A_{2n}[g]
  =
  -a_{2n}E_{2n}
  +
  \sum_\alpha c_\alpha I_\alpha
  +
  \nabla_\mu J^\mu .
  \label{eq:weyl-anomaly-type-abd-general}
\end{equation}
Here
\begin{equation}
  E_{2n}
  =
  \frac{1}{2^n}
  \delta^{\mu_1\nu_1\cdots\mu_n\nu_n}
         _{\rho_1\sigma_1\cdots\rho_n\sigma_n}
  R^{\rho_1\sigma_1}{}_{\mu_1\nu_1}\cdots
  R^{\rho_n\sigma_n}{}_{\mu_n\nu_n}
  \label{eq:euler-density-general}
\end{equation}
is the Euler density in the curvature convention used in this chapter.  The
term \(E_{2n}\) is the type-A anomaly.  Its integral is topological:
\(\int_M\sqrt g\,E_{2n}\) is proportional to the Euler characteristic on a
closed \(2n\)-manifold.  Consequently its Weyl variation is a total
derivative, so it obeys
\eqref{eq:weyl-wess-zumino-consistency}.  The scalar densities \(I_\alpha\)
are type-B anomalies: \(\sqrt g\,I_\alpha\) is locally Weyl invariant in
dimension \(2n\).  The divergence \(\nabla_\mu J^\mu\) denotes type-D, or
trivial, terms; these are Weyl variations of local metric counterterms and
are scheme dependent.

This classification is local.  It classifies the contact terms in
stress-tensor Ward identities produced by varying a smooth background metric.
Global terms, boundaries, defects, and parity-odd orientation-dependent
candidates require additional data and are not part of the parity-even bulk
list in \eqref{eq:weyl-anomaly-type-abd-general}.  In odd dimension there is
no parity-even local bulk trace anomaly of this type on a closed smooth
manifold, although finite nonlocal data such as the sphere free energy can be
important conformal data.
\end{quotedtheorem}

In two dimensions, with the convention above, write
\begin{equation}
  \mathcal A_2[g]=-\frac{c_{2d}}{24\pi}R .
  \label{eq:two-dimensional-weyl-anomaly}
\end{equation}
This is purely type A: \(E_2=R\), and there is no independent parity-even
type-B scalar in two dimensions.

\begin{proposition}[Four-dimensional parity-even Weyl anomaly]
\label{prop:four-dimensional-parity-even-weyl-anomaly}
In four dimensions, define
\[
  E_4
  =
  R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma}
  -4R_{\mu\nu}R^{\mu\nu}
  +R^2,
  \qquad
  W^2
  =
  W_{\mu\nu\rho\sigma}W^{\mu\nu\rho\sigma},
\]
where \(W_{\mu\nu\rho\sigma}\) is the Weyl tensor.  A parity-even CFT has
\begin{equation}
  \mathcal A_4[g]
  =
  \frac{1}{16\pi^2}
  \left(
    c_{\rm W} W^2
    -a_{\rm W} E_4
    +b_{\rm D}\nabla^2 R
  \right).
  \label{eq:four-dimensional-weyl-anomaly}
\end{equation}
The coefficient \(b_{\rm D}\) is changed by adding a local \(R^2\) counterterm
to \(\mathcal W[g]\).  The coefficients \(a_{\rm W}\) and \(c_{\rm W}\) are
unchanged by such counterterms at a conformal fixed point.
\(E_4\) is the type-A term, \(W^2\) is the unique parity-even type-B scalar,
and \(\nabla^2R\) is type D.  More explicitly,
\[
  \delta_\omega\int_M\dd^4x\,\sqrt g\,R^2
  =
  -12\int_M\dd^4x\,\sqrt g\,\omega\,\nabla^2R
\]
on a closed four-manifold, so the coefficient of \(\nabla^2R\) is shifted by
a choice of local \(R^2\) counterterm.
The completeness of this parity-even four-dimensional list follows from the
dimension-four curvature basis
\(R_{\mu\nu\rho\sigma}^2\), \(R_{\mu\nu}^2\), \(R^2\), and \(\nabla^2R\):
the two nontrivial Wess--Zumino classes are the Euler combination \(E_4\) and
the Weyl-invariant combination \(W^2\), while the remaining scalar can be
changed by the \(R^2\) counterterm just displayed.
\end{proposition}

\begin{proof}
The independent parity-even dimension-four local scalar densities are spanned,
up to total derivatives and algebraic curvature identities, by
\(R_{\mu\nu\rho\sigma}^2\), \(R_{\mu\nu}^2\), \(R^2\), and \(\nabla^2R\).
The two cohomologically nontrivial combinations solving
\eqref{eq:weyl-wess-zumino-consistency} are the Euler density
\(E_4\) and the Weyl-invariant density \(W^2\).  The displayed Weyl variation
of \(\int\sqrt g\,R^2\) shows that \(\nabla^2R\) is a counterterm variation and
therefore type D.  This proves the stated local classification in the
parity-even metric sector.
\end{proof}

\begin{quotedtheorem}[Six-dimensional parity-even Weyl anomaly]
\label{qthm:six-dimensional-parity-even-weyl-anomaly}
In six dimensions, modulo type-D total-derivative terms, the parity-even bulk
classification has one type-A coefficient and three type-B coefficients:
\begin{equation}
  \mathcal A_6[g]
  =
  \frac{1}{(4\pi)^3}
  \left(
    -a_6E_6+c_1I_1+c_2I_2+c_3I_3
    +\nabla_\mu J^\mu
  \right).
  \label{eq:six-dimensional-weyl-anomaly-classification}
\end{equation}
A convenient representative of the three type-B invariants is
\begin{align}
  I_1
  &=
  W_{\mu\rho\sigma\nu}
  W^{\mu\alpha\beta\nu}
  W_{\alpha}{}^{\rho\sigma}{}_{\beta},
  \notag\\
  I_2
  &=
  W_{\mu\nu}{}^{\rho\sigma}
  W_{\rho\sigma}{}^{\alpha\beta}
  W_{\alpha\beta}{}^{\mu\nu},
  \notag\\
  I_3
  &\equiv
  W_{\mu\nu\rho\sigma}
  \left(
    \nabla^2\delta^\mu{}_\alpha
    +4R^\mu{}_\alpha
    -\frac65R\,\delta^\mu{}_\alpha
  \right)
  W^{\alpha\nu\rho\sigma}
  \quad
  \bmod \nabla_\mu J^\mu .
  \label{eq:six-dimensional-type-b-invariants}
\end{align}
The equivalence sign in the definition of \(I_3\) is essential: different
representatives differ by type-D terms, while the coefficient \(c_3\) of the
cohomology class is invariant under local counterterm changes.  In dimensions
higher than six the same rule applies, but the number of independent Weyl
invariants grows because more contractions of Weyl tensors and conformally
covariant derivatives are available.
For \(D=6\), dimension counting and the algebraic Bianchi identities leave
two independent cubic Weyl contractions and one independent
two-Weyl-tensor invariant with two derivatives; Ricci-dependent representatives
either combine with these into the conformally covariant operator in \(I_3\),
enter the Euler density \(E_6\), or are type-D counterterm variations.
\end{quotedtheorem}

\section{Anomaly Coefficients and Stress-Tensor Correlators}

\begin{convention}[\(2D\) central charge and anomaly]
\label{conv:two-dimensional-central-charge-weyl-anomaly}
When a two-dimensional CFT is coupled to a curved metric and its flat
holomorphic stress tensor is normalized by the same metric source functional,
the symbol \(c_{2d}\) denotes both the coefficient in
\eqref{eq:two-dimensional-weyl-anomaly} and the Virasoro central charge in the
holomorphic OPE normalization
\[
  \langle T(z)T(0)\rangle=\frac{c_{2d}}{2z^4}.
\]
The identification is the contact Ward identity obtained by taking two metric
variations of \eqref{eq:two-dimensional-weyl-anomaly} around a flat metric and
then rewriting the result in complex coordinates.  The present chapter uses
this identity as a normalization convention; it is not being introduced here
as a separate theorem without proof.
\end{convention}

\begin{convention}[Four-dimensional \(C_T\) and \(c_{\rm W}\)]
\label{conv:four-dimensional-ct-cw-normalization}
Use the following local normalization of the separated flat-space
stress-tensor two-point function:
\[
  \left\langle T_{\mu\nu}(x)T_{\rho\sigma}(0)\right\rangle
  =
  \frac{C_T}{|x|^{2D}}
  I_{\mu\nu,\rho\sigma}(x),
\]
where
\[
  I_{\mu\nu,\rho\sigma}(x)
  =
  \frac12\bigl(
    I_{\mu\rho}(x)I_{\nu\sigma}(x)
    +
    I_{\mu\sigma}(x)I_{\nu\rho}(x)
  \bigr)
  -
  \frac1D\delta_{\mu\nu}\delta_{\rho\sigma}.
\]
The same two-point convention is developed systematically in
Chapter~\ref{ch:volume-iii-correlation-functions-conformal-frames}.  With
this normalization, the symbol \(c_{\rm W}\) in
\eqref{eq:four-dimensional-weyl-anomaly} is normalized so that the separated
stress-tensor two-point coefficient is
\begin{equation}
  C_T=\frac{40}{\pi^4}\,c_{\rm W}.
  \label{eq:four-dimensional-ct-cw-relation}
\end{equation}
The relation is a normalization statement for the metric source.  Its
curvature-kernel verification is obtained by taking two metric variations of
the \(W^2\) term in \eqref{eq:four-dimensional-weyl-anomaly}, Fourier
transforming the resulting quadratic curvature contact term, and matching the
logarithmic scale dependence of the flat-space \(TT\) correlator to the
displayed two-point normalization.  The Euler term begins to contribute
independent information at the next metric variation.
\end{convention}

\begin{remark}[Four-dimensional anomaly map from \(TTT\) data]
\label{rem:four-dimensional-anomaly-map-from-ttt}
Use the local convention in which the parity-even separated \(TTT\) correlator
in \(D=4\) is expanded in three Osborn--Petkou structures,
\[
  \langle T_{\mu\nu}(x_1)T_{\rho\sigma}(x_2)T_{\alpha\beta}(x_3)\rangle_{\rm even}
  =
  a_{\rm OP}\mathcal I^{(1)}
  +b_{\rm OP}\mathcal I^{(2)}
  +c_{\rm OP}\mathcal I^{(3)} .
\]
Chapter~\ref{ch:volume-iii-correlation-functions-conformal-frames} constructs
these three separated tensor structures explicitly from conformal covariance,
conservation, and permutation symmetry.  The small-sphere stress-tensor Ward
identity gives, in this normalization and at \(D=4\),
\begin{equation}
  C_T
  =
  \frac{\pi^2}{3}
  \left(
    14a_{\rm OP}-2b_{\rm OP}-5c_{\rm OP}
  \right),
  \label{eq:op-coefficients-ct-four-d}
\end{equation}
so \(c_{\rm W}\) is this same linear combination through
\eqref{eq:four-dimensional-ct-cw-relation}.  The Euler coefficient
\(a_{\rm W}\) is a second, independent linear functional of
\((a_{\rm OP},b_{\rm OP},c_{\rm OP})\), obtained by taking three metric
variations of the \(E_4\) term in
\eqref{eq:four-dimensional-weyl-anomaly} and projecting the resulting trace
Ward identity onto the three separated tensor structures.  Equivalently, one
may use the free-field basis of four-dimensional stress-tensor structures:
\begin{equation}
  \langle TTT\rangle
  =
  n_s\langle TTT\rangle_{\rm real\ scalar}
  +
  n_f\langle TTT\rangle_{\rm Dirac\ fermion}
  +
  n_v\langle TTT\rangle_{\rm Maxwell\ vector}.
  \label{eq:ttt-free-field-basis}
\end{equation}
In that basis,
\begin{equation}
  a_{\rm W}
  =
  \frac{n_s}{360}
  +
  \frac{11n_f}{360}
  +
  \frac{31n_v}{180},
  \qquad
  c_{\rm W}
  =
  \frac{n_s}{120}
  +
  \frac{n_f}{20}
  +
  \frac{n_v}{10}.
  \label{eq:four-dimensional-free-basis-anomaly-map}
\end{equation}
Here \(n_f\) multiplies the Dirac-fermion stress-tensor structure; using a
Weyl fermion instead halves the fermion entries.  The third independent
linear combination of \(a_{\rm OP},b_{\rm OP},c_{\rm OP}\) is genuine \(TTT\)
data not fixed by \(a_{\rm W}\) and \(c_{\rm W}\).  After \(C_T\) has fixed
the overall stress-tensor normalization, Chapter~\ref{ch:cft-light-ray-energy-correlators}
uses the conformal-collider coordinates \(t_2,t_4\) as detector coordinates on
the remaining separated \(TTT\) data.
\end{remark}

\section{Improvement of the Stress Tensor}

\paragraph{Scalar improvement of the flat stress tensor.}
The stress tensor has improvement freedom.  If \(L(x)\) is a local scalar
operator, then
\begin{equation}
  T_{\mu\nu}
  \longmapsto
  T'_{\mu\nu}
  =
  T_{\mu\nu}
  +
  \left(
  \delta_{\mu\nu}\partial^2-\partial_\mu\partial_\nu
  \right)L
  \label{eq:scalar-stress-improvement}
\end{equation}
preserves conservation in flat space.  More general improvements are possible,
but \eqref{eq:scalar-stress-improvement} is the case needed for the free
scalar and many scalar fixed points.  Its trace is
\[
  T'^\mu{}_\mu
  =
  T^\mu{}_\mu
  +(D-1)\partial^2L .
\]
If \(T^\mu{}_\mu\) is a total Laplacian of a local scalar, an improvement can
make the trace vanish in flat space.
The added tensor is identically conserved in flat space because
\[
  \partial^\mu(\delta_{\mu\nu}\partial^2-\partial_\mu\partial_\nu)L
  =
  \partial_\nu\partial^2L-\partial^2\partial_\nu L
  =
  0 .
\]
Taking its trace gives
\((D-1)\partial^2L\), which is the displayed trace shift.

The improvement can be represented by adding a curvature coupling to the
background-metric action.  For a scalar field \(\phi\), the term
\[
  \Delta S_\xi
  =
  \frac{\xi}{2}
  \int_M\dd^D x\,\sqrt g\,R(g)\phi^2
\]
vanishes in flat space when \(R(g)=0\), but its metric variation contributes
to \(T_{\mu\nu}\) at flat \(g_{\mu\nu}\).  Thus a term absent from the
flat-space action can change the flat-space stress tensor.

\section{The Conformally Coupled Free Scalar}

\begin{proposition}[Conformally coupled free scalar]
\label{prop:conformally-coupled-free-scalar}
For a real massless scalar field in \(D>2\) Euclidean dimensions, the curvature
coupling
\[
  \xi_\ast=\frac{D-2}{4(D-1)}
\]
is the unique scalar-curvature improvement of the form
\(\frac{\xi}{2}\int\sqrt g\,R\phi^2\) for which the flat-space stress tensor
has vanishing trace modulo the scalar equation of motion.  The corresponding
curved action is \eqref{eq:conformal-scalar-curved-action}.
\end{proposition}

\begin{proof}
Consider a real massless scalar field in Euclidean signature.  The minimally
coupled curved-space action is
\[
  S_0[\phi;g]
  =
  \frac12\int_M\dd^D x\,\sqrt g\,
  g^{\mu\nu}\partial_\mu\phi\,\partial_\nu\phi .
\]
Its metric variation is
\[
  \delta_g S_0
  =
  \frac12\int_M\dd^Dx\,\sqrt g\,
  \delta g^{\mu\nu}
  \left(
  \partial_\mu\phi\,\partial_\nu\phi
  -
  \frac12g_{\mu\nu}g^{\rho\sigma}
  \partial_\rho\phi\,\partial_\sigma\phi
  \right).
\]
On the flat background this gives
\[
  T_{\mu\nu}^{(0)}
  =
  \partial_\mu\phi\,\partial_\nu\phi
  -
  \frac12\delta_{\mu\nu}(\partial\phi)^2 .
\]
Its trace is
\[
  T^{(0)\mu}{}_\mu
  =
  \left(1-\frac D2\right)(\partial\phi)^2 .
\]
For \(D>2\) this trace is a nonzero local operator before improvement.  Using
the equation of motion \(\partial^2\phi=0\), the identity
\[
  \partial^2\phi^2
  =
  2(\partial\phi)^2+2\phi\,\partial^2\phi
\]
shows that the trace is a Laplacian modulo the equation of motion.

Now add the curvature functional
\[
  S_1[\phi;g]
  =
  \frac12\int_M\dd^Dx\,\sqrt g\,R(g)\phi^2 .
\]
The scalar curvature variation with respect to \(g^{\mu\nu}\) is
\[
  \delta R
  =
  R_{\mu\nu}\delta g^{\mu\nu}
  +
  \nabla_\rho
  \left(
  g^{\mu\nu}\delta\Gamma^\rho_{\mu\nu}
  -
  g^{\mu\rho}\delta\Gamma^\nu_{\mu\nu}
  \right),
\]
where \(\delta\Gamma^\rho_{\mu\nu}\) is the variation of the Levi-Civita
connection.  Integrating the total derivative by parts and then setting
\(g_{\mu\nu}=\delta_{\mu\nu}\) gives
\[
  \delta_g S_1\big|_{g=\delta}
  =
  \frac12\int_{\R^D}\dd^Dx\,\delta g^{\mu\nu}
  \left(
  -\partial_\mu\partial_\nu\phi^2
  +
  \delta_{\mu\nu}\partial^2\phi^2
  \right).
\]
The remaining terms in the full curved-space variation are proportional to
curvature and vanish after the metric is set to \(\delta_{\mu\nu}\).

For
\[
  S[\phi;g]=S_0[\phi;g]+aS_1[\phi;g],
\]
the flat-space stress tensor is
\begin{equation}
  T_{\mu\nu}
  =
  \partial_\mu\phi\,\partial_\nu\phi
  -
  \frac12\delta_{\mu\nu}(\partial\phi)^2
  +
  a
  \left(
  \delta_{\mu\nu}\partial^2-\partial_\mu\partial_\nu
  \right)\phi^2,
  \label{eq:improved-free-scalar-stress}
\end{equation}
The total translation generators
\(P^\nu=\int_\Sigma\dd^{D-1}x\,T^{0\nu}\) are unchanged by the improvement
term when spatial total derivatives integrate to zero on the constant-time
slice.  The trace changes:
\[
  T^\mu{}_\mu
  =
  \left(1-\frac D2\right)(\partial\phi)^2
  +
  a(D-1)\partial^2\phi^2
\]
and therefore
\[
  T^\mu{}_\mu
  =
  \left(1-\frac D2+2(D-1)a\right)(\partial\phi)^2
  +
  2(D-1)a\,\phi\,\partial^2\phi .
\]
The trace vanishes modulo the scalar equation of motion if and only if
\begin{equation}
  a=\xi_\ast=\frac{D-2}{4(D-1)}.
  \label{eq:conformal-scalar-xi}
\end{equation}
The corresponding curved-space action is
\begin{equation}
  S[\phi;g]
  =
  \frac12
  \int_M\dd^D x\,\sqrt g\,
  \left[
  g^{\mu\nu}\partial_\mu\phi\partial_\nu\phi
  +
  \xi_\ast R(g)\phi^2
  \right].
  \label{eq:conformal-scalar-curved-action}
\end{equation}
With the Weyl transformation
\[
  g_{\mu\nu}\mapsto \ee^{2\omega}g_{\mu\nu},
  \qquad
  \phi\mapsto \ee^{-\frac{D-2}{2}\omega}\phi,
\]
this action is Weyl invariant at the classical level by
Proposition~\ref{prop:conformal-laplacian-weyl-covariance}.  In the
corresponding free massless scalar QFT, the improved trace is
\[
  \widehat T^\mu{}_\mu(x)
  =
  2(D-1)\xi_\ast\,\widehat\phi(x)\,\partial^2\widehat\phi(x).
\]
It represents the zero class in the local-operator vector space modulo the
free equation-of-motion ideal.  Equivalently, for local insertions
\(\mathcal O_i(y_i)\) with \(x\ne y_i\),
\[
  \left\langle
    \widehat T^\mu{}_\mu(x)\,
    \prod_i\mathcal O_i(y_i)
  \right\rangle
  =
  0 .
\]
At coincident points the factor \(\partial^2\widehat\phi\) generates the
Schwinger--Dyson contact terms of the free Gaussian measure.
\end{proof}

\begin{proposition}[Weyl covariance of the conformal Laplacian]
\label{prop:conformal-laplacian-weyl-covariance}
Let \(D>2\), set
\[
  \alpha=\frac{D-2}{2},
  \qquad
  \xi_\ast=\frac{D-2}{4(D-1)},
\]
and define the conformal Laplacian
\[
  P_g=-\nabla_g^2+\xi_\ast R(g).
\]
Under a Weyl transformation
\[
  g'_{\mu\nu}=\ee^{2\omega}g_{\mu\nu},
  \qquad
  \phi'=\ee^{-\alpha\omega}\phi,
\]
one has
\begin{equation}
  P_{g'}\phi'
  =
  \ee^{-\frac{D+2}{2}\omega}P_g\phi .
  \label{eq:conformal-laplacian-weyl-covariance}
\end{equation}
Consequently, on a closed manifold or with boundary conditions eliminating
boundary terms,
\[
  \frac12\int_M\dd^D x\,\sqrt g\,
  \phi\,P_g\phi
  =
  \frac12\int_M\dd^D x\,\sqrt{g'}\,
  \phi'\,P_{g'}\phi' .
\]
This is precisely the Weyl invariance of
\eqref{eq:conformal-scalar-curved-action}.
\end{proposition}

\begin{proof}
For a scalar \(f\), the Laplacian transforms as
\[
  \nabla_{g'}^2f
  =
  \ee^{-2\omega}
  \left(
    \nabla_g^2f
    +(D-2)\nabla^\mu\omega\,\nabla_\mu f
  \right),
\]
where covariant derivatives and contractions on the right use \(g\).  Apply
this to \(f=\ee^{-\alpha\omega}\phi\).  Since \(2\alpha=D-2\), the first
derivative terms in \(\phi\) cancel:
\[
  \nabla_{g'}^2(\ee^{-\alpha\omega}\phi)
  =
  \ee^{-(\alpha+2)\omega}
  \left(
    \nabla_g^2\phi
    -\alpha(\nabla_g^2\omega)\phi
    -\alpha^2(\nabla\omega)^2\phi
  \right).
\]
The scalar curvature transforms by
\[
  R(g')
  =
  \ee^{-2\omega}
  \left(
    R(g)-2(D-1)\nabla_g^2\omega
    -(D-1)(D-2)(\nabla\omega)^2
  \right).
\]
Therefore
\[
\begin{aligned}
  P_{g'}(\ee^{-\alpha\omega}\phi)
  &=
  \ee^{-(\alpha+2)\omega}
  \bigl[
    -\nabla_g^2\phi
    +\xi_\ast R(g)\phi
\\
  &\qquad
    +(\alpha-2(D-1)\xi_\ast)(\nabla_g^2\omega)\phi
\\
  &\qquad
    +(\alpha^2-(D-1)(D-2)\xi_\ast)(\nabla\omega)^2\phi
  \bigr].
\end{aligned}
\]
The two coefficients involving \(\omega\) vanish exactly for
\(\xi_\ast=(D-2)/(4(D-1))\).  Since \(\alpha+2=(D+2)/2\), this proves
\eqref{eq:conformal-laplacian-weyl-covariance}.  Finally,
\(\sqrt{g'}=\ee^{D\omega}\sqrt g\), and
\[
  D-\alpha-\frac{D+2}{2}=0.
\]
The integrated quadratic form is therefore invariant.  Integrating
\(\phi(-\nabla_g^2)\phi\) by parts rewrites the quadratic form as
\eqref{eq:conformal-scalar-curved-action}.
\end{proof}

Figure~\ref{fig:stress-improvement-curvature} records the dimension-dependent
coefficient in the conformal Laplacian.  The displayed values are a useful
check on the normalization: in \(D=4\) the coefficient is \(1/6\), while
\(D=2\) is degenerate because the free scalar has \(\Delta_\phi=0\) and the
curvature improvement coefficient vanishes.

\begin{figure}[H]
  \centering
  \resizebox{0.84\textwidth}{!}{%
  \renewcommand{\arraystretch}{1.24}
  \begin{tabular}{|c|c|c|c|}
    \hline
    \(D\) & \(\Delta_\phi=(D-2)/2\) &
    \(\xi_\ast=(D-2)/[4(D-1)]\) &
    conformal Laplacian \(P_g\) \\
    \hline
    \(2\) & \(0\) & \(0\) & \(-\nabla_g^2\) \\
    \hline
    \(3\) & \(1/2\) & \(1/8\) &
    \(-\nabla_g^2+\frac18R(g)\) \\
    \hline
    \(4\) & \(1\) & \(1/6\) &
    \(-\nabla_g^2+\frac16R(g)\) \\
    \hline
    \(5\) & \(3/2\) & \(3/16\) &
    \(-\nabla_g^2+\frac{3}{16}R(g)\) \\
    \hline
    \(6\) & \(2\) & \(1/5\) &
    \(-\nabla_g^2+\frac15R(g)\) \\
    \hline
    \(10\) & \(4\) & \(2/9\) &
    \(-\nabla_g^2+\frac29R(g)\) \\
    \hline
  \end{tabular}
  }
  \caption{The conformal improvement coefficient is fixed by Weyl covariance
  of \(P_g=-\nabla_g^2+\xi_\ast R(g)\):
  \(P_{\ee^{2\omega}g}(\ee^{-(D-2)\omega/2}\phi)
  =\ee^{-(D+2)\omega/2}P_g\phi\).  Varying the metric-coupled action with
  this \(P_g\), and only then setting \(g_{\mu\nu}=\delta_{\mu\nu}\), gives the
  improved flat-space stress tensor.}
  \label{fig:stress-improvement-curvature}
\end{figure}

\section{Ward Identities with Stress-Tensor Insertions}

\begin{proposition}[Stress-tensor Ward identity with spinning insertions]
\label{prop:stress-ward-identity-spinning-insertions}
Let
\[
  X=
  \mathcal O_{A_1}(x_1)\cdots\mathcal O_{A_n}(x_n),
  \qquad
  X_i[B]=
  \mathcal O_{A_1}(x_1)\cdots B(x_i)\cdots
  \mathcal O_{A_n}(x_n),
\]
where \(B(x_i)\) replaces the \(i\)-th operator.  Suppose first that each
\(\mathcal O_{A_i}\) belongs to a finite-dimensional source representation of
linearized diffeomorphisms, with matrices
\((\mathsf R^{\mu\nu}_{(i)})_{A_i}{}^{B_i}\) defined by the linear term
\[
  \delta_\xi \mathcal O_{A_i}(x_i)
  =
  \xi^\rho(x_i)\partial_\rho\mathcal O_{A_i}(x_i)
  +
  \partial_\mu\xi_\nu(x_i)
  (\mathsf R^{\mu\nu}_{(i)})_{A_i}{}^{B_i}
  \mathcal O_{B_i}(x_i).
\]
For scalar operators the matrices \(\mathsf R^{\mu\nu}_{(i)}\) vanish.  For
operators carrying only orthonormal-frame spin, the antisymmetric part of
\(\mathsf R^{\mu\nu}_{(i)}\) is the usual spin matrix
\(\mathsf S^{\mu\nu}_{(i)}=-\mathsf S^{\nu\mu}_{(i)}\), and the symmetric
traceless part is absent.  The source-functional identity
\eqref{eq:cft-source-diffeomorphism-ward}, differentiated with respect to the
corresponding sources and then evaluated at zero source, gives the smeared
flat-space Ward identity
\begin{equation}
  \int\dd^D x\,
  \partial_\mu\xi_\nu(x)
  \left\langle T^{\mu\nu}(x)X\right\rangle
  =
  \sum_{i=1}^n
  \xi^\nu(x_i)\partial_{x_i^\nu}\langle X\rangle
  +
  \sum_{i=1}^n
  \partial_\mu\xi_\nu(x_i)
  \left\langle
    X_i[
    (\mathsf R^{\mu\nu}_{(i)})_{A_i}{}^{B_i}\mathcal O_{B_i}]
  \right\rangle .
  \label{eq:stress-ward-spinning-smeared}
\end{equation}
Equivalently, as an identity of distributions in \(x\),
\begin{equation}
  \partial_\mu
  \left\langle
  T^{\mu\nu}(x)X
  \right\rangle
  =
  -\sum_{i=1}^n
  \delta(x-x_i)
  \partial_{x_i^\nu}
  \langle X\rangle
  +
  \sum_{i=1}^n
  \partial_\mu
  \left[
    \delta(x-x_i)
    \left\langle
      X_i[
      (\mathsf R^{\mu\nu}_{(i)})_{A_i}{}^{B_i}\mathcal O_{B_i}]
    \right\rangle
  \right].
  \label{eq:stress-ward-spinning-local}
\end{equation}
There is no omitted spin term in
\eqref{eq:stress-ward-spinning-local}.  For scalar insertions
\(\mathsf R^{\mu\nu}_{(i)}=0\), and the familiar translation Ward identity is
obtained by deleting the second line.  For spinning insertions that second
line is part of the identity.  To see its relation to the conformal Ward
identities used later, let the \(i\)-th operator be a primary in an
orthonormal-frame spin representation \(\rho_i\), with infinitesimal spin
matrices \(S^{(i)}_{\mu\nu}\) in the convention of
Chapter~\ref{ch:volume-iii-primary-operators-finite-transformations}.  If
\(\epsilon^\mu\) is a conformal Killing vector and
\[
  J^\mu_\epsilon=T^\mu{}_\nu\epsilon^\nu,
  \qquad
  \sigma_\epsilon=\frac1D\partial_\rho\epsilon^\rho,
  \qquad
  \omega_{\epsilon,\mu\nu}
  =\frac12(\partial_\nu\epsilon_\mu-\partial_\mu\epsilon_\nu),
\]
then smearing \eqref{eq:stress-ward-spinning-smeared} with
\(\xi=\epsilon\) and adding the trace Ward identity gives
\begin{equation}
  \partial_\mu
  \left\langle J^\mu_\epsilon(x)X\right\rangle
  =
  \sum_{i=1}^n
  \delta^{(D)}(x-x_i)
  \left\langle
    X_i[\mathcal D^{(i)}_\epsilon\mathcal O_{A_i}]
  \right\rangle
  +
  \sigma_\epsilon(x)\mathcal A(x;X),
  \label{eq:stress-ward-primary-spin-current}
\end{equation}
where
\begin{equation}
  \mathcal D^{(i)}_\epsilon\mathcal O_{i,a_i}(x_i)
  =
  -\epsilon^\rho(x_i)\partial_{x_i^\rho}\mathcal O_{i,a_i}(x_i)
  -\Delta_i\sigma_\epsilon(x_i)\mathcal O_{i,a_i}(x_i)
  -\frac{\ii}{2}\omega_\epsilon^{\mu\nu}(x_i)
  (S^{(i)}_{\mu\nu})_{a_i}{}^{b_i}
  \mathcal O_{i,b_i}(x_i).
  \label{eq:stress-ward-primary-spin-contact-operator}
\end{equation}
Equations \eqref{eq:stress-ward-primary-spin-current} and
\eqref{eq:stress-ward-primary-spin-contact-operator} are the same source
Ward identity as \eqref{eq:primary-spin-current-ward}; the derivative
contact term in \eqref{eq:stress-ward-spinning-local} is precisely what
becomes the local rotation term proportional to
\(\omega_\epsilon^{\mu\nu}S^{(i)}_{\mu\nu}\).  Thus one may use the scalar
translation formula only after verifying that all insertions are scalars, or
after the spin contact term has been carried along separately.
For a basis of composite operators that mixes under the linearized
diffeomorphism, the displayed representation matrix is replaced by the
corresponding source-chart representation matrix.  The trace Ward identity
similarly records the response under Weyl transformations.  At a flat-space
conformal fixed point, the separated-point trace insertion vanishes, while
contact terms encode scaling dimensions and conformal transformations of the
inserted operators.

The charge form of the Ward identity follows by integrating
\(T^\mu{}_\nu\epsilon^\nu\) over a small sphere around each insertion.  This
is the bridge between the local stress-tensor equation and the conformal
algebra acting on local operators.
\end{proposition}
