\chapter{Correlation Functions and Conformal Frames}
\label{ch:volume-iii-correlation-functions-conformal-frames}
\ConventionsEuclideanCFT


\section{Ward Identities for Correlators}

Throughout this chapter, correlators are Euclidean vacuum correlators at
separated insertion points.  The symbol \(\ket{\vac}\) denotes the common
vacuum identified in
Section~\ref{sec:flat-conformal-cylinder-vacua}: the Poincare vacuum of the
flat Wightman theory, the conformally invariant vacuum under the hypotheses of
Proposition~\ref{prop:poincare-vacuum-is-conformal-vacuum}, and the
zero-energy cylinder ground state after the Weyl-map normalization of radial
quantization.  With this identification, and with the special-conformal
vacuum invariance derived in
Lemma~\ref{lem:vacuum-k-invariance-from-radial-spectrum}, the global Ward
identity of Chapter~\ref{ch:volume-iii-conformal-charges-ward-identities}
applies.  If
\(\epsilon^\mu(x)\partial_\mu\) is a conformal Killing vector and
\(\delta_\epsilon\mathcal O_i\) is the corresponding infinitesimal variation of
the \(i\)th local operator, then
\begin{equation}
  \sum_{i=1}^n
  \left\langle
    \mathcal O_1(x_1)\cdots
    \delta_\epsilon\mathcal O_i(x_i)\cdots
    \mathcal O_n(x_n)
  \right\rangle
  =
  0
  \label{eq:correlator-global-ward-identity}
\end{equation}
as a distribution on the configuration space \(x_i\ne x_j\).  Contact terms on
coincident-point diagonals are part of the local Ward identities and are not
used in the kinematic classification below.

\begin{definition}[Separated Euclidean CFT correlator]
\label{def:separated-euclidean-cft-correlator}
For an ordered \(n\)-tuple of local operators
\(\mathcal O_1,\ldots,\mathcal O_n\), the separated configuration space is
\[
  \operatorname{Conf}_n(\R^D)
  =
  \{(x_1,\ldots,x_n)\in(\R^D)^n\mid x_i\ne x_j\text{ for }i\ne j\}.
\]
The separated correlator is the restriction of the corresponding
Schwinger distribution to this open set,
\[
  G_{\mathcal O_1\cdots\mathcal O_n}(x_1,\ldots,x_n)
  =
  \left\langle
    \mathcal O_1(x_1)\cdots\mathcal O_n(x_n)
  \right\rangle,
\]
with values in the dual tensor product of the spin representation spaces when
the insertions carry indices.  A scalar primary tuple
\((\mathcal O_i,\Delta_i)\) is separated conformally covariant if, for every
finite conformal transformation \(f\) whose image avoids coincidences,
\[
  G_{\mathcal O_1\cdots\mathcal O_n}(f(x_1),\ldots,f(x_n))
  =
  \prod_{i=1}^n \Omega_f(x_i)^{-\Delta_i}\,
  G_{\mathcal O_1\cdots\mathcal O_n}(x_1,\ldots,x_n),
  \label{eq:separated-scalar-correlator-covariance}
\]
where \(f^*\delta=\Omega_f^2\delta\).  This definition contains no extension
across the collision diagonals.  Such extensions are local renormalization
data and can differ by contact terms supported on partial diagonals.
\end{definition}

\begin{lemma}[Special-conformal invariance from radial spectrum positivity]
\label{lem:vacuum-k-invariance-from-radial-spectrum}
Let \(\ket{\vac}\) be the common vacuum of
Section~\ref{sec:flat-conformal-cylinder-vacua}.  Assume that
\(\widehat D_{\rm rad}\) is a nonnegative self-adjoint generator of cylinder
time translations, that
\[
  \widehat D_{\rm rad}\ket{\vac}=0,
\]
and that the special conformal generators are defined on the vacuum and obey,
on the relevant common domain,
\begin{equation}
  [\widehat D_{\rm rad},\widehat K_\mu]=-\widehat K_\mu .
  \label{eq:radial-d-k-vacuum-commutator}
\end{equation}
Then
\[
  \widehat K_\mu\ket{\vac}=0 .
\]
Thus the special-conformal part of the vacuum Ward identity follows from
radial spectral positivity and the conformal commutation relation; it is not an
extra invariance assumption.
\end{lemma}

\begin{proof}
Let \(\psi_\mu=\widehat K_\mu\ket{\vac}\).  By the domain hypothesis and
\eqref{eq:radial-d-k-vacuum-commutator},
\[
  \widehat D_{\rm rad}\psi_\mu
  =
  \widehat D_{\rm rad}\widehat K_\mu\ket{\vac}
  =
  \widehat K_\mu\widehat D_{\rm rad}\ket{\vac}
  -
  \widehat K_\mu\ket{\vac}
  =
  -\psi_\mu .
\]
If \(\psi_\mu\ne0\), this would be an eigenvector of the nonnegative
self-adjoint operator \(\widehat D_{\rm rad}\) with eigenvalue \(-1\), which is
impossible.  Hence \(\psi_\mu=0\).
\end{proof}

Equivalently, for a finite conformal transformation \(f:x\mapsto \widetilde x\)
implemented by a unitary operator \(U_f\), invariance of the vacuum gives
\[
  \left\langle
    \mathcal O_1(x_1)\cdots\mathcal O_n(x_n)
  \right\rangle
  =
  \left\langle
    (U_f\mathcal O_1U_f^{-1})(x_1)\cdots
    (U_f\mathcal O_nU_f^{-1})(x_n)
  \right\rangle .
\]
The operators on the right are written at the same coordinate labels as before;
their transformation laws supply the Jacobian and spin factors.  For a scalar
primary \(\phi\) of dimension \(\Delta_\phi\),
\begin{equation}
  \widetilde\phi(\widetilde x)
  =
  \left|\det\frac{\partial\widetilde x}{\partial x}\right|^{-\Delta_\phi/D}
  \phi(x).
  \label{eq:scalar-primary-finite-transform-correlators}
\end{equation}

\section{One- and Two-Point Functions}

\begin{proposition}[Scalar one-point selection rule]
\label{prop:scalar-one-point-selection-rule}
Let \(\mathcal O_i\) be scalar primary operators of dimensions
\(\Delta_i\).  If the separated Ward identities and the finite scalar
primary covariance law hold, then
\[
  \left\langle \mathcal O_i(x)\right\rangle=0
  \qquad\text{unless}\qquad
  \Delta_i=0.
\]
In a reflection-positive CFT whose \(\Delta=0\) scalar primary sector is the
vacuum representation, the only nonzero scalar-primary one-point function is
the identity expectation value, after choosing the identity normalization.
\end{proposition}

\begin{proof}
Translation invariance makes a one-point function constant:
\[
  \left\langle \mathcal O_i(x)\right\rangle=c_i.
\]
Scale covariance gives
\[
  c_i=\lambda^{-\Delta_i}c_i
  \qquad
  \text{for all }\lambda>0.
\]
If \(\Delta_i\ne0\), choosing any \(\lambda\ne1\) forces \(c_i=0\).  The final
statement uses the \(\Delta=0\) branch of the scalar unitarity analysis in
Chapter~\ref{ch:volume-iii-unitarity-bounds-short-multiplets}: the local field
has vanishing translation descendants and hence acts as a multiple of the
identity on the cyclic vacuum representation.
\end{proof}

\begin{theorem}[Scalar two-point selection rule]
\label{thm:scalar-two-point-selection-rule}
Assume the separated-point global conformal Ward identities
\eqref{eq:correlator-global-ward-identity} and scalar primary transformation
law \eqref{eq:scalar-primary-finite-transform-correlators}.  If
\(\mathcal O_i,\mathcal O_j\) are scalar primaries of dimensions
\(\Delta_i,\Delta_j\), then
\begin{equation}
  \left\langle \mathcal O_i(x)\mathcal O_j(y)\right\rangle
  =
  \begin{cases}
    \displaystyle \frac{C_{ij}}{|x-y|^{2\Delta_i}},
    &\Delta_i=\Delta_j,\\[1.0em]
    0,&\Delta_i\ne\Delta_j .
  \end{cases}
  \label{eq:scalar-primary-two-point}
\end{equation}
\end{theorem}

\begin{proof}
Translation and rotation invariance imply
\[
  \left\langle \mathcal O_i(x)\mathcal O_j(y)\right\rangle
  =
  F_{ij}(|x-y|).
\]
Scale covariance gives
\[
  F_{ij}(r)=\frac{C_{ij}}{r^{\Delta_i+\Delta_j}}.
\]
The condition \(\Delta_i=\Delta_j\) for a nonzero coefficient follows from a
conformal transformation that fixes both points but has different local scale
factors at the two points.  Put the two points at \(0\) and \(e_1\), where
\(e_1\) is a fixed unit vector, and define
\[
  f_\lambda(x)^\mu
  =
  \lambda
  \frac{x^\mu+a^\mu x^2}{1+2a\cdot x+a^2x^2},
  \qquad
  a^\mu=(\lambda-1)e_1^\mu,
  \qquad
  \lambda>0 .
\]
Then \(f_\lambda(0)=0\), \(f_\lambda(e_1)=e_1\),
\[
  \left.\frac{\partial f_\lambda^\mu}{\partial x^\nu}\right|_{x=0}
  =
  \lambda\,\delta^\mu{}_\nu,
  \qquad
  \left.\frac{\partial f_\lambda^\mu}{\partial x^\nu}\right|_{x=e_1}
  =
  \lambda^{-1}\delta^\mu{}_\nu .
\]
Applying \eqref{eq:scalar-primary-finite-transform-correlators} to
\(\langle\mathcal O_i(e_1)\mathcal O_j(0)\rangle\) gives
\[
  C_{ij}=\lambda^{\Delta_i-\Delta_j}C_{ij}
  \qquad\text{for all }\lambda>0,
\]
so \(C_{ij}=0\) unless \(\Delta_i=\Delta_j\).  Substituting this condition in
the scale-covariant form gives \eqref{eq:scalar-primary-two-point}.
\end{proof}
Within a finite-dimensional space of primaries with the same dimension and the
same spin, one may choose a basis that diagonalizes the Hermitian matrix
\(C_{ij}\).

\section{Radial Inner Product and BPZ Conjugation}

Let \(x=rn\), \(r>0\), \(n\in S^{D-1}\), and set
\(\tau=\log r\).  The Weyl map from \(\R^D\setminus\{0\}\) to
\(\R_\tau\times S^{D-1}\) relates a scalar primary on flat space to the
cylinder operator by
\begin{equation}
  \mathcal O(x)=r^{-\Delta}\widetilde{\mathcal O}(\tau,n).
  \label{eq:flat-cylinder-scalar-primary}
\end{equation}
The state created by insertion at the origin is
\begin{equation}
  \ket{\mathcal O}
  =
  \lim_{\tau\to-\infty}
  \ee^{-\Delta\tau}
  \widetilde{\mathcal O}(\tau,n)\ket{\vac},
  \label{eq:ket-from-cylinder-limit}
\end{equation}
where \(\ket{\vac}\) is the common vacuum specified above.  For a primary the
limit is independent of \(n\).  The conjugate bra is obtained by the opposite
cylinder-time limit,
\begin{equation}
  \langle\!\langle \mathcal O|
  =
  \lim_{\tau\to+\infty}
  \ee^{\Delta\tau}
  \bra{\vac}\widetilde{\mathcal O}(\tau,n).
  \label{eq:bpz-bra-from-cylinder-limit}
\end{equation}
This radial conjugation is often called BPZ conjugation.  In a
reflection-positive theory, after including the ordinary internal adjoint of the
local operator, it is the Hilbert-space adjoint of the state.  Consequently the
two-point coefficient has the inner-product interpretation
\begin{equation}
  \langle\!\langle \mathcal O_i|\mathcal O_j\rangle
  =
  \lim_{\tau_i\to+\infty,\ \tau_j\to-\infty}
  \ee^{\Delta_i\tau_i-\Delta_j\tau_j}
  \left\langle
    \widetilde{\mathcal O}_i(\tau_i,n_i)
    \widetilde{\mathcal O}_j(\tau_j,n_j)
  \right\rangle
  =
  C_{ij},
  \label{eq:bpz-inner-product-two-point-coefficient}
\end{equation}
for primaries with equal dimension.  Positivity of this matrix is the same
positivity used in the unitarity-bound derivation.

\begin{figure}[H]
  \centering
  \resizebox{0.90\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      insertion/.style={circle,fill=violet!80!black,inner sep=1.8pt},
      arrow/.style={-{Latex[length=2mm]},line width=0.8pt},
      cylinder/.style={line width=0.85pt}
    }
    \begin{scope}
      \node[font=\small] at (0,2.05) {flat space};
      \draw[thick] (-2.1,-1.3) rectangle (2.1,1.35);
      \node[insertion] at (-0.75,-0.2) {};
      \node[violet!80!black,below left] at (-0.75,-0.2) {\(\mathcal O_j(0)\)};
      \node[insertion] at (0.95,0.45) {};
      \node[violet!80!black,above right] at (0.95,0.45) {\(\mathcal O_i(x)\)};
      \draw[blue!65!black,line width=0.8pt] (-0.75,-0.2)--(0.95,0.45);
      \node[blue!65!black] at (0.05,0.48) {\(|x|\)};
    \end{scope}

    \draw[arrow] (2.65,0.05)--(4.00,0.05);
    \node[align=center] at (3.33,0.58) {radial\\quantization};

    \begin{scope}[shift={(6.35,0)}]
      \node[font=\small] at (0,2.05) {cylinder};
      \draw[cylinder] (-1.15,1.25) ellipse (1.05 and 0.25);
      \draw[cylinder] (-2.20,1.25)--(-2.20,-1.25);
      \draw[cylinder] (-0.10,1.25)--(-0.10,-1.25);
      \draw[cylinder] (-1.15,-1.25) ellipse (1.05 and 0.25);
      \node[violet!80!black] at (-1.15,-0.70) {\(\ket{\mathcal O_j}\)};
      \node[violet!80!black] at (-1.15,0.78)
        {\(\langle\!\langle\mathcal O_i|\)};
      \draw[green!45!black,line width=0.8pt] (-2.20,0.15) arc[start angle=180,end angle=360,x radius=1.05,y radius=0.25];
      \draw[green!45!black,dashed,line width=0.8pt] (-2.20,0.15) arc[start angle=180,end angle=0,x radius=1.05,y radius=0.25];
      \draw[arrow,orange!85!black] (0.38,-1.10)--(0.38,1.10);
      \node[orange!85!black,right] at (0.38,0.05) {\(\tau\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The scalar two-point coefficient is the radial-quantization inner
  product between the state created at the origin and the BPZ-conjugate
  insertion at future cylinder time.}
  \label{fig:two-point-cylinder-inner-product}
\end{figure}

\section{Spin Structures}

\begin{definition}[Inversion tensor and symmetric-traceless transport]
\label{def:inversion-stt-parallel-transport}
The inversion tensor
\begin{equation}
  I_{\mu\nu}(x)
  =
  \delta_{\mu\nu}-2\frac{x_\mu x_\nu}{x^2}
  \label{eq:inversion-tensor}
\end{equation}
is the orthogonal part of the inversion Jacobian.  It is the basic
building block for two-point functions with vector or tensor indices.  Its
action on the line spanned by \(x^\mu\) is multiplication by \(-1\), while its
action on the orthogonal complement of \(x^\mu\) is the identity.  Hence
\(\det I(x)=-1\): inversion lies in \(O(D)\setminus SO(D)\).  Every use of
\(I_{\mu\nu}(x)\) in a spin tensor structure therefore presupposes the extension
of the \(SO(D)\) spin action to the disconnected orthogonal, or pin, action
specified in Chapter~\ref{ch:volume-iii-primary-operators-finite-transformations}.

For a rank-\(s\) symmetric traceless tensor representation, define
\[
  \mathcal I^{(s)}_{\mu_1\cdots\mu_s,\nu_1\cdots\nu_s}(x)
  =
  \Pi_{\operatorname{STT}}
  \left[
    I_{\mu_1\nu_1}(x)\cdots I_{\mu_s\nu_s}(x)
  \right],
\]
where \(\Pi_{\operatorname{STT}}\) symmetrizes separately in the
\(\mu\)-indices and the \(\nu\)-indices and removes all traces with respect to
\(\delta_{\mu\nu}\).
\end{definition}

\begin{proposition}[Spinning two-point structures]
\label{prop:spinning-two-point-structures}
Let \(V_\mu\) be a vector primary of dimension \(\Delta_V\).  After
diagonalizing degeneracy labels within the same primary type,
\begin{equation}
  \left\langle V_\mu(x)V_\nu(0)\right\rangle
  =
  \frac{C_V}{|x|^{2\Delta_V}}I_{\mu\nu}(x).
  \label{eq:vector-primary-two-point}
\end{equation}
More generally, if
\(\mathcal O_{\mu_1\cdots\mu_s}\) is a rank-\(s\) symmetric traceless tensor
primary of dimension \(\Delta_{\mathcal O}\), then
\begin{equation}
  \left\langle
    \mathcal O_{\mu_1\cdots\mu_s}(x)
    \mathcal O_{\nu_1\cdots\nu_s}(0)
  \right\rangle
  =
  \frac{C_{\mathcal O}}{|x|^{2\Delta_{\mathcal O}}}
  \mathcal I^{(s)}_{\mu_1\cdots\mu_s,\nu_1\cdots\nu_s}(x),
  \label{eq:stt-primary-two-point}
\end{equation}
up to the same positive Hermitian coefficient matrix in possible degeneracy
labels.  Two-point functions between inequivalent \(SO(D)\) representations
vanish by rotational covariance.
\end{proposition}

\begin{proof}
Translation covariance puts one insertion at the origin, and dilatation
covariance gives the overall power \(|x|^{-2\Delta}\) once the two
dimensions have been matched by the scalar selection rule generalized to the
chosen spin type.  The remaining tensor must intertwine the spin
representation at the origin with the spin representation transported to the
point \(x\).  Applying inversion sends \(x\) to \(x/x^2\), fixes the origin as
the point at infinity in radial quantization, and contributes the orthogonal
Jacobian \(I_{\mu\nu}(x)\) on vector indices.  Hence the unique vector
intertwiner is \(I_{\mu\nu}(x)\).  For symmetric traceless rank \(s\), the
tensor product of \(s\) vector intertwiners must be projected to the
symmetric-traceless irreducible component at both ends, giving
\(\mathcal I^{(s)}\).  Schur's lemma on each irreducible \(SO(D)\) type gives
vanishing between inequivalent representations and a Hermitian degeneracy
matrix within equivalent representations.  Reflection positivity makes that
matrix positive semidefinite before quotienting null states and positive
definite on the Hilbert quotient.
\end{proof}

For the stress tensor, \(s=2\), \(\Delta_T=D\), and conservation gives a
canonical normalization convention through the Ward identity.  Its two-point
function is
\begin{equation}
  \left\langle T_{\mu\nu}(x)T_{\rho\sigma}(0)\right\rangle
  =
  \frac{C_T}{|x|^{2D}}
  I_{\mu\nu,\rho\sigma}(x),
  \label{eq:stress-tensor-two-point}
\end{equation}
where
\[
  I_{\mu\nu,\rho\sigma}(x)
  =
  \frac12\bigl(
    I_{\mu\rho}(x)I_{\nu\sigma}(x)
    +
    I_{\mu\sigma}(x)I_{\nu\rho}(x)
  \bigr)
  -
  \frac1D\delta_{\mu\nu}\delta_{\rho\sigma}.
\]
The coefficient \(C_T\) is part of the intrinsic CFT data because the
normalization of \(T_{\mu\nu}\) is fixed by the translation Ward identity.

\section{Three-Point Functions}

\begin{theorem}[Scalar three-point kinematics]
\label{thm:scalar-three-point-kinematics}
Let \(\mathcal O_i\) be scalar primaries of dimensions \(\Delta_i\).  On the
configuration space of three distinct points, separated conformal covariance
fixes the three-point function to the form
\begin{equation}
  \left\langle
    \mathcal O_1(x_1)\mathcal O_2(x_2)\mathcal O_3(x_3)
  \right\rangle
  =
  \frac{C_{123}}
  {|x_{12}|^{\Delta_1+\Delta_2-\Delta_3}
   |x_{23}|^{\Delta_2+\Delta_3-\Delta_1}
   |x_{13}|^{\Delta_1+\Delta_3-\Delta_2}},
  \label{eq:scalar-primary-three-point}
\end{equation}
where
\[
  x_{ij}=x_i-x_j.
\]
\end{theorem}

\begin{proof}
The connected Euclidean conformal group acts transitively on ordered triples
of distinct points.  Indeed, translate \(x_3\) to the origin, use an inversion
composed with a translation to send one of the remaining points to infinity,
and then rotate and dilate the last finite point to a fixed unit vector.  Thus
there is no conformal cross-ratio for three separated points; the correlator
is determined by one coefficient once its local weights at the three
insertions are fixed.

It remains to determine the powers of the pairwise distances.  Write the
unique possible separated scalar form as
\[
  C_{123}
  |x_{12}|^{-\alpha_{12}}
  |x_{23}|^{-\alpha_{23}}
  |x_{13}|^{-\alpha_{13}} .
\]
Under inversion \(x\mapsto x/x^2\),
\[
  |x_{ij}|
  \longmapsto
  \frac{|x_{ij}|}{|x_i|\,|x_j|},
  \qquad
  \Omega(x_i)=|x_i|^{-2}.
\]
Comparison with
\eqref{eq:separated-scalar-correlator-covariance} gives the three local
weight equations
\[
  \alpha_{12}+\alpha_{13}=2\Delta_1,\qquad
  \alpha_{12}+\alpha_{23}=2\Delta_2,\qquad
  \alpha_{13}+\alpha_{23}=2\Delta_3 .
\]
Solving this linear system gives
\[
  \alpha_{12}=\Delta_1+\Delta_2-\Delta_3,\quad
  \alpha_{23}=\Delta_2+\Delta_3-\Delta_1,\quad
  \alpha_{13}=\Delta_1+\Delta_3-\Delta_2,
\]
which is precisely \eqref{eq:scalar-primary-three-point}.  Uniform dilatation
covariance is then automatic because the sum of the three powers is
\(\Delta_1+\Delta_2+\Delta_3\).
\end{proof}

The constants \(C_{123}\), after fixing the two-point normalization of
primaries, are the scalar three-point OPE coefficients.  If the primaries
carry spin, let \(a_i\) denote indices in the corresponding finite-dimensional
\(SO(D)\) representations.  Conformal covariance says that
\begin{equation}
  \left\langle
    \mathcal O_{1,a_1}(x_1)
    \mathcal O_{2,a_2}(x_2)
    \mathcal O_{3,a_3}(x_3)
  \right\rangle
  =
  \sum_{r=1}^{n_{123}}
  C_{123}^{(r)}
  \mathcal I_{a_1a_2a_3}^{(r)}(x_{12},x_{13}),
  \label{eq:spinning-three-point-structures}
\end{equation}
where the tensors \(\mathcal I^{(r)}\) form a basis of conformally covariant
three-point tensor structures, including the required homogeneity in the
separations.  The finite integer \(n_{123}\) is representation-theoretic; the
numbers \(C_{123}^{(r)}\) are dynamical data after a convention for two-point
normalization has been fixed.

The stress tensor illustrates why some three-point coefficients are
constrained but not removed from the data.  In a parity-even CFT in
\(D\ge4\), the separated three-point function of a symmetric conserved
traceless stress tensor is classified by the following Osborn--Petkou
basis.\footnote{This is the position-space basis of H.~Osborn and
A.~C.~Petkou, \emph{Implications of Conformal Invariance in Field Theories
for General Dimensions}, Ann. Phys. 231 (1994), 311--362,
arXiv:hep-th/9307010, specialized here to the separated-point, parity-even
stress-tensor three-point function and rewritten in the conventions of this
chapter.}  Set
\[
  X\equiv X_{12,3}
  =
  \frac{x_{13}}{x_{13}^2}
  -
  \frac{x_{23}}{x_{23}^2},
  \qquad
  \widehat X_\mu=\frac{X_\mu}{(X^2)^{1/2}},
\]
and, for a dimensionless tensor
\(t_{\mu\nu,\rho\sigma,\alpha\beta}(\widehat X)\), define
\begin{equation}
  \mathfrak C[t]_{\mu\nu,\rho\sigma,\alpha\beta}
  =
  \frac{
    I_{\mu\nu,\mu'\nu'}(x_{13})
    I_{\rho\sigma,\rho'\sigma'}(x_{23})
    t_{\mu'\nu',\rho'\sigma',\alpha\beta}(\widehat X)}
  {|x_{12}|^D |x_{13}|^D |x_{23}|^D}.
  \label{eq:op-ttt-prefactor}
\end{equation}
Let \(\operatorname{Sym}_{\mu\nu}\) denote the unnormalized symmetrization
obtained by adding the expression with \(\mu\) and \(\nu\) interchanged.  The
following symmetric-traceless tensors are
\begin{align}
  h^1_{\mu\nu}
  &=
  \widehat X_\mu\widehat X_\nu-\frac1D\delta_{\mu\nu},
  \nonumber\\
  h^2_{\mu\nu,\rho\sigma}
  &=
  \widehat X_\mu\widehat X_\rho\delta_{\nu\sigma}
  +\widehat X_\nu\widehat X_\rho\delta_{\mu\sigma}
  +\widehat X_\mu\widehat X_\sigma\delta_{\nu\rho}
  +\widehat X_\nu\widehat X_\sigma\delta_{\mu\rho}
  \nonumber\\
  &\quad
  -\frac4D
  \left(
    \widehat X_\mu\widehat X_\nu\delta_{\rho\sigma}
    +
    \widehat X_\rho\widehat X_\sigma\delta_{\mu\nu}
  \right)
  +
  \frac4{D^2}\delta_{\mu\nu}\delta_{\rho\sigma},
  \nonumber\\
  h^3_{\mu\nu,\rho\sigma}
  &=
  \delta_{\mu\rho}\delta_{\nu\sigma}
  +
  \delta_{\mu\sigma}\delta_{\nu\rho}
  -
  \frac2D\delta_{\mu\nu}\delta_{\rho\sigma},
  \nonumber\\
  h^4_{\mu\nu,\rho\sigma,\alpha\beta}
  &=
  \operatorname{Sym}_{\rho\sigma}
  \operatorname{Sym}_{\alpha\beta}
  \left[
    h^3_{\mu\nu,\rho\alpha}
    \widehat X_\sigma\widehat X_\beta
  \right]
  -
  \frac2D\delta_{\rho\sigma}h^2_{\mu\nu,\alpha\beta}
  \nonumber\\
  &\quad
  -
  \frac2D\delta_{\alpha\beta}h^2_{\mu\nu,\rho\sigma}
  -
  \frac8{D^2}\delta_{\rho\sigma}\delta_{\alpha\beta}h^1_{\mu\nu},
  \nonumber\\
  h^5_{\mu\nu,\rho\sigma,\alpha\beta}
  &=
  \operatorname{Sym}_{\mu\nu}
  \operatorname{Sym}_{\rho\sigma}
  \operatorname{Sym}_{\alpha\beta}
  \left[
    \delta_{\mu\rho}\delta_{\nu\alpha}\delta_{\sigma\beta}
  \right]
  \nonumber\\
  &\quad
  -
  \frac4D
  \left(
    \delta_{\mu\nu}h^3_{\rho\sigma,\alpha\beta}
    +
    \delta_{\rho\sigma}h^3_{\mu\nu,\alpha\beta}
    +
    \delta_{\alpha\beta}h^3_{\mu\nu,\rho\sigma}
  \right)
  -
  \frac8{D^2}
  \delta_{\mu\nu}\delta_{\rho\sigma}\delta_{\alpha\beta}.
  \label{eq:op-h-tensors}
\end{align}
Each \(h^r\) is symmetric and traceless in every displayed index pair.  The
most general parity-even separated-point structure satisfying stress-tensor
conservation is the three-parameter tensor
\begin{align}
  t[a,b,c]_{\mu\nu,\rho\sigma,\alpha\beta}
  &=
  a\,h^5_{\mu\nu,\rho\sigma,\alpha\beta}
  +
  b\,h^4_{\alpha\beta,\mu\nu,\rho\sigma}
  -
  (2a+b)
  \left(
    h^4_{\mu\nu,\rho\sigma,\alpha\beta}
    +
    h^4_{\rho\sigma,\mu\nu,\alpha\beta}
  \right)
  \nonumber\\
  &\quad
  +
  c\,h^3_{\mu\nu,\rho\sigma}h^1_{\alpha\beta}
  +
  c
  \left(
    h^3_{\rho\sigma,\alpha\beta}h^1_{\mu\nu}
    +
    h^3_{\mu\nu,\alpha\beta}h^1_{\rho\sigma}
  \right)
  \nonumber\\
  &\quad
  +
  e\,h^2_{\mu\nu,\rho\sigma}h^1_{\alpha\beta}
  +
  e'
  \left(
    h^2_{\rho\sigma,\alpha\beta}h^1_{\mu\nu}
    +
    h^2_{\mu\nu,\alpha\beta}h^1_{\rho\sigma}
  \right)
  \nonumber\\
  &\quad
  +
  f\,h^1_{\mu\nu}h^1_{\rho\sigma}h^1_{\alpha\beta},
  \label{eq:op-ttt-tensor}
\end{align}
where conservation fixes the auxiliary coefficients to be
\begin{align}
  e=(D+2)(Da+b-c),
  \qquad
  e'=-(D+4)(D-2)a-(D-2)b+Dc,
  \nonumber\\
  f=(D+4)(D-2)(4a+2b-c).
  \label{eq:op-ttt-dependent-coefficients}
\end{align}
The three tensor structures in
\eqref{eq:stress-tensor-three-point-structures} are therefore
\[
  \mathcal I^{(1)}=\mathfrak C[t[1,0,0]],
  \qquad
  \mathcal I^{(2)}=\mathfrak C[t[0,1,0]],
  \qquad
  \mathcal I^{(3)}=\mathfrak C[t[0,0,1]].
\]
Equivalently,
\begin{equation}
  \left\langle
    T_{\mu\nu}(x_1)T_{\rho\sigma}(x_2)T_{\alpha\beta}(x_3)
  \right\rangle
  =
  a\,\mathcal I^{(1)}_{\mu\nu,\rho\sigma,\alpha\beta}
  +
  b\,\mathcal I^{(2)}_{\mu\nu,\rho\sigma,\alpha\beta}
  +
  c\,\mathcal I^{(3)}_{\mu\nu,\rho\sigma,\alpha\beta}
  =
  \mathfrak C[t[a,b,c]]_{\mu\nu,\rho\sigma,\alpha\beta},
  \label{eq:stress-tensor-three-point-structures}
\end{equation}
with three independent coefficients \(a,b,c\) when \(D\ge4\).  The
independence statement is not a counting by inspection.  Before imposing
conservation, conformal covariance, tracelessness, and the exchange symmetries
of the three stress tensors reduce the parity-even separated-point ansatz to
the five coefficients \(a,b,c,e,f\); the coefficient \(e'\) appearing in
\eqref{eq:op-ttt-tensor} is then fixed by
\[
  e+e'=8a+4b-2c .
\]
The conservation equation
\[
  \left(\partial_\mu-D\frac{X_\mu}{X^2}\right)
  t_{\mu\nu,\rho\sigma,\alpha\beta}(X)=0
\]
imposes two independent linear equations.  In the parametrization above these
equations are exactly the formulae
\eqref{eq:op-ttt-dependent-coefficients}, which determine
\(e,e'\), and \(f\) in terms of \(a,b,c\).  To see that no further relation is
hidden for \(D\ge4\), put the three insertion points in a collinear frame by a
conformal transformation.  The residual group is \(O(D-1)\).  The completely
transverse part of the conserved tensor is then parametrized by three
\(O(D-1)\)-invariant tensors with coefficients \(r,s,u\), and the
Osborn--Petkou comparison gives
\begin{equation}
  D^2 r=16a-2Db+(D+4)c,
  \qquad
  D s=-4a-c,
  \qquad
  u=a .
  \label{eq:op-collinear-rsu}
\end{equation}
For \(D\ge4\), the transverse space has dimension at least three, and these
three invariant tensors are linearly independent.  Hence
\((a,b,c)\mapsto(r,s,u)\) has no universal null direction in the space of
separated-point parity-even \(TTT\) structures.

For \(D=3\), the transverse space in the collinear frame is two-dimensional.
\footnote{For the three-dimensional stress-tensor/current structure count see
J.~Maldacena and A.~Zhiboedov, \emph{A Note on CFT Correlators in Three
Dimensions}, arXiv:1104.4317.  For the general spinning-correlator counting
technology and its later conformal-block implementation, see M.~S.~Costa,
J.~Penedones, D.~Poland, and S.~Rychkov, \emph{Spinning Conformal
Correlators}, arXiv:1107.3554, and M.~S.~Costa, T.~Hansen, J.~Penedones, and
E.~Trevisani, \emph{Radial Expansion for Spinning Conformal Blocks},
arXiv:1603.05552.}
The three transverse invariant tensors obey one Schouten identity, and the
collinear tensor depends only on
\begin{equation}
  r-4u,
  \qquad
  s+2u .
  \label{eq:op-d3-two-invariants}
\end{equation}
Using \eqref{eq:op-collinear-rsu} at \(D=3\), the null direction is
\[
  (a,b,c)=(1,-1,2).
\]
Thus, as separated parity-even three-dimensional structures,
\begin{equation}
  \mathcal I^{(1)}_{D=3}
  -
  \mathcal I^{(2)}_{D=3}
  +
  2\mathcal I^{(3)}_{D=3}
  =
  0,
  \qquad
  \mathcal I^{(3)}_{D=3}
  =
  \frac12\left(
    \mathcal I^{(2)}_{D=3}-\mathcal I^{(1)}_{D=3}
  \right).
  \label{eq:op-d3-ttt-relation}
\end{equation}
Consequently a parity-even three-dimensional stress-tensor three-point
function has two separated structures, for instance
\[
  \left\langle TTT\right\rangle_{D=3}^{\rm even}
  =
  A\,\mathcal I^{(1)}_{D=3}
  +
  B\,\mathcal I^{(2)}_{D=3}.
  \label{eq:op-d3-two-ttt-structures}
\]
If parity is not imposed in \(D=3\), there is a further parity-odd separated
structure; it is not part of the parity-even Osborn--Petkou basis displayed
here.

The Ward identity fixes one
linear combination of these coefficients in terms of the two-point coefficient
\(C_T\).  Let
\[
  S_D=\frac{2\pi^{D/2}}{\Gamma(D/2)}
\]
be the area of the unit \((D-1)\)-sphere.  The small-sphere Ward identity uses
the angular averages
\[
  \int_{S^{D-1}}\dd\Omega\,\widehat n_\mu\widehat n_\nu
  =
  \frac{S_D}{D}\delta_{\mu\nu},
  \qquad
  \int_{S^{D-1}}\dd\Omega\,
  \widehat n_\mu\widehat n_\nu\widehat n_\rho\widehat n_\sigma
  =
  \frac{S_D}{D(D+2)}
  \left(
    \delta_{\mu\nu}\delta_{\rho\sigma}
    +
    \delta_{\mu\rho}\delta_{\nu\sigma}
    +
    \delta_{\mu\sigma}\delta_{\nu\rho}
  \right).
\]
Substituting \eqref{eq:op-ttt-tensor} into the flux
\(\int_{\partial B_\epsilon(x_3)}\dd S\,n_\lambda
T^{\lambda\tau}\epsilon_\tau\), taking \(\epsilon\to0\), and comparing with
the conformal variation of the two remaining stress tensors in
\eqref{eq:stress-tensor-two-point} gives
\begin{equation}
  4S_D\,
  \frac{(D-2)(D+3)a-2b-(D+1)c}{D(D+2)}
  =
  C_T .
  \label{eq:ttt-ward-ct-relation}
\end{equation}
The remaining two independent combinations are genuine CFT data.  In lower
dimensions or in parity-violating theories the list of admissible structures
changes; the statement above is tied to the stated assumptions.

\begin{figure}[H]
  \centering
  \resizebox{0.82\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      insertion/.style={circle,fill=violet!80!black,inner sep=1.8pt},
      sphere/.style={blue!65!black,line width=0.85pt},
      arrow/.style={-{Latex[length=2mm]},line width=0.8pt}
    }
    \node[insertion] at (-3.1,0.75) {};
    \node[violet!80!black,above] at (-3.1,0.75) {\(T_{\mu\nu}(x_1)\)};
    \node[insertion] at (3.1,0.75) {};
    \node[violet!80!black,above] at (3.1,0.75) {\(T_{\rho\sigma}(x_2)\)};
    \node[insertion] at (0,-0.35) {};
    \node[violet!80!black,below] at (0,-0.35) {\(T_{\alpha\beta}(x_3)\)};
    \draw[sphere] (0,-0.35) circle (0.95);
    \draw[arrow,blue!65!black] (0,-0.35)--(0.68,0.32);
    \node[blue!65!black,right] at (0.68,0.32) {\(n^\alpha\)};
    \draw[orange!85!black,line width=0.8pt]
      (-3.1,0.75) .. controls (-1.85,1.45) and (-0.85,1.05) .. (0,0.60)
      .. controls (0.85,1.05) and (1.85,1.45) .. (3.1,0.75);
    \node[align=center] at (0,-1.85)
      {\(\displaystyle
        Q_\epsilon(S_{x_3})=
        \int_{S_{x_3}}\dd\sigma\,
        n^\alpha T_{\alpha\beta}\epsilon^\beta
      \)};
    \draw[arrow] (4.1,-0.35)--(5.35,-0.35);
    \node[align=center] at (6.85,-0.35)
      {acts on the other\\stress tensors by the\\conformal Ward identity};
  \end{tikzpicture}
  }
  \caption{The Ward identity relates a small-sphere stress-tensor flux inside
  \(\langle TTT\rangle\) to the conformal variation of the remaining
  insertions.  This fixes the Ward-normalized part of the \(TTT\) data in
  terms of \(C_T\).}
  \label{fig:stress-tensor-three-point-ward}
\end{figure}

As an important mixed example, the Ward identity fixes the coefficient of the
stress-tensor three-point function with two identical scalar primaries.  For
\(\Delta_1=\Delta_2=\Delta\),
\begin{equation}
  \left\langle
    \mathcal O(x_1)\mathcal O(x_2)T_{\mu\nu}(x_3)
  \right\rangle
  =
  \frac{C_{\mathcal O\mathcal O T}}
  {|x_{13}|^D|x_{23}|^D|x_{12}|^{2\Delta-D}}
  \left(
    \frac{X_\mu X_\nu}{X^2}
    -
    \frac1D\delta_{\mu\nu}
  \right),
  \label{eq:scalar-scalar-stress-tensor-three-point}
\end{equation}
with
\[
  X^\mu=\frac{x_{13}^\mu}{x_{13}^2}
       -\frac{x_{23}^\mu}{x_{23}^2}.
\]
More generally, for two scalar primaries \(\mathcal O_1,\mathcal O_2\) of the
same dimension \(\Delta\) with
\[
  \left\langle
    \mathcal O_1(x)\mathcal O_2(0)
  \right\rangle
  =
  \frac{N_{12}}{|x|^{2\Delta}},
\]
the same tensor structure appears with coefficient
\begin{equation}
  C_{12T}
  =
  -\,\frac{D\Delta}{(D-1)A_{D-1}}\,N_{12},
  \qquad
  A_{D-1}=\operatorname{vol}(S^{D-1})
  =
  \frac{2\pi^{D/2}}{\Gamma(D/2)}.
  \label{eq:scalar-scalar-stress-ward-coefficient}
\end{equation}
Here the sign is the sign of the stress-tensor convention in
\eqref{eq:cft-charge-surface} and \eqref{eq:variation-by-small-charge}.  The
derivation is a small-sphere Ward identity for the dilatation vector
\(\epsilon^\nu=x^\nu\).  Put \(x_1=x\), \(x_2=0\), \(x_3=y=r n\), where
\(|n|=1\), and send \(r\to0\) with \(x\) fixed.  Then
\[
  X^\mu
  =
  \frac{(x-y)^\mu}{|x-y|^2}
  -
  \frac{(-y)^\mu}{|y|^2}
  =
  \frac{n^\mu}{r}+\frac{x^\mu}{|x|^2}+O\!\left(\frac{r}{|x|^2}\right),
\]
and therefore
\[
  \frac{X_\mu X_\nu}{X^2}-\frac1D\delta_{\mu\nu}
  =
  n_\mu n_\nu-\frac1D\delta_{\mu\nu}+O\!\left(\frac{r}{|x|}\right),
  \qquad
  \frac1{|x_{13}|^D|x_{23}|^D|x_{12}|^{2\Delta-D}}
  =
  \frac{r^{-D}}{|x|^{2\Delta}}
  \left(1+O\!\left(\frac{r}{|x|}\right)\right).
\]
The sphere \(S_2\) is oriented as the boundary of the punctured integration
domain, hence \(\dd S_\mu=-r^{D-1}n_\mu\dd\Omega\), while
\(\epsilon_\nu(y)=r n_\nu\).  The leading flux is consequently
\[
  \left\langle
    \mathcal O_1(x)Q_\epsilon(S_2)\mathcal O_2(0)
  \right\rangle
  =
  -\,\frac{C_{12T}}{|x|^{2\Delta}}
  \int_{S^{D-1}}\dd\Omega\,
  n^\mu n^\nu
  \left(
    n_\mu n_\nu-\frac1D\delta_{\mu\nu}
  \right),
\]
up to terms that vanish as \(r\to0\).  The angular projection is
\begin{align}
  \int_{S^{D-1}}\dd\Omega\,
  n^\mu n^\nu
  \left(
    n_\mu n_\nu-\frac1D\delta_{\mu\nu}
  \right)
  &=
  \int_{S^{D-1}}\dd\Omega\,
  \left[
    (n^\mu n_\mu)(n^\nu n_\nu)
    -
    \frac1D n^\mu n_\mu
  \right]
  \nonumber\\
  &=
  \int_{S^{D-1}}\dd\Omega\,
  \left[
    1-\frac1D
  \right]
  =
  A_{D-1}\left(1-\frac1D\right).
  \label{eq:scalar-stress-ward-angular-projection}
\end{align}
In the second line we used \(n^\mu n_\mu=1\) pointwise on the unit sphere;
this is the entire reduction of the fourth-order angular expression in this
particular contraction.
so the flux equals
\[
  -\,\frac{D-1}{D}A_{D-1}
  \frac{C_{12T}}{|x|^{2\Delta}}.
\]
The same orientation convention converts the local dilatation Ward identity
\(\delta_\epsilon\mathcal O_2(0)=-\Delta\mathcal O_2(0)\) into
\[
  \left\langle
    \mathcal O_1(x)Q_\epsilon(S_2)\mathcal O_2(0)
  \right\rangle
  =
  \frac{\Delta N_{12}}{|x|^{2\Delta}}.
\]
Matching the two expressions gives
\eqref{eq:scalar-scalar-stress-ward-coefficient}.  If
\(\Delta_1\ne\Delta_2\), the two-point coefficient \(N_{12}\) vanishes by
dilatation covariance, and the corresponding stress-tensor three-point
function vanishes.

\section{Four-Point Functions and Cross Ratios}

\begin{definition}[Ordered four-point conformal frame]
\label{def:ordered-four-point-conformal-frame}
For \(D\ge2\), let
\[
  x=(x_1,x_2,x_3,x_4)\in\operatorname{Conf}_4(\R^D).
\]
A four-point conformal frame for \(x\) is a conformal transformation, possibly
followed by the one-point compactification of \(\R^D\), that sends the
quadruple to
\[
  z_1=0,\qquad z_2=z,\qquad z_3=1,\qquad z_4=\infty
\]
inside an oriented two-plane.  The associated scalar cross ratios are
\begin{equation}
  u=
  \frac{x_{12}^2x_{34}^2}{x_{13}^2x_{24}^2},
  \qquad
  v=
  \frac{x_{14}^2x_{23}^2}{x_{13}^2x_{24}^2}.
  \label{eq:cross-ratios-u-v}
\end{equation}
In the frame they obey
\[
  z=\frac{z_{12}z_{34}}{z_{13}z_{24}},
  \qquad
  u=z\bar z,\qquad v=(1-z)(1-\bar z).
\]
\end{definition}

\begin{theorem}[Completeness of four-point cross ratios]
\label{thm:four-point-cross-ratio-completeness}
For a generic ordered quadruple of distinct Euclidean points in \(D\ge2\), an
oriented four-point frame is labelled by \(z\).  The pair \((u,v)\) is
equivalent to \(z\) modulo complex conjugation.  For \(D\ge3\), this pair
separates connected conformal orbits.  For \(D=2\), it separates orbits only
after adjoining the orientation-reversing conformal transformation
\(z\mapsto\bar z\); the connected orientation-preserving group distinguishes
\(z\) from \(\bar z\).  Thus scalar four-point correlators invariant under
the relevant Euclidean parity operation can have nontrivial position
dependence only through two independent real conformal invariants.
\end{theorem}

\begin{proof}
We give the explicit frame construction.  Translate so that \(x_1=0\).  The
special conformal map
\[
  y\longmapsto
  \frac{y-b\,y^2}{1-2b\cdot y+b^2y^2},
  \qquad
  b=\frac{x_4}{x_4^2},
\]
fixes the origin and sends \(x_4\) to the point at infinity, because the
denominator vanishes at \(y=x_4\).  For a generic quadruple, the image of
\(x_3\) is finite and nonzero; a rotation and a dilatation send it to the unit
vector \(e_1\).  The stabilizer of \(0\), \(e_1\), and \(\infty\) is the
orthogonal group acting on the \((D-1)\)-dimensional subspace orthogonal to
\(e_1\).  This stabilizer rotates the image of \(x_2\) into the
\((e_1,e_2)\)-plane, leaving a complex coordinate \(z\).  In this frame the
Euclidean distances give exactly
\[
  u=z\bar z,\qquad v=(1-z)(1-\bar z).
\]
Conversely, two framed quadruples with the same \(z\) are equal, and replacing
the orientation of the two-plane sends \(z\) to \(\bar z\) without changing
\((u,v)\).  When \(D\ge3\), this orientation reversal of the chosen two-plane
is implemented by a rotation in the orthogonal complement, for instance by
flipping one in-plane basis vector and one transverse basis vector.  When
\(D=2\), it is not in the connected orientation-preserving conformal group and
must be added as a separate parity operation.  The dimension count is
consistent:
\[
  4D-\frac{(D+1)(D+2)}2+\frac{(D-2)(D-3)}2=2,
\]
where the last term is the dimension of the residual rotation group fixing a
generic two-plane frame.
\end{proof}

\begin{figure}[H]
  \centering
  \resizebox{0.94\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      insertion/.style={circle,fill=violet!80!black,inner sep=1.8pt},
      arrow/.style={-{Latex[length=2mm]},line width=0.8pt}
    }
    \begin{scope}
      \node[font=\small] at (0,2.0) {four points in \(\R^D\)};
      \draw[thick] (-2.25,-1.35) rectangle (2.25,1.35);
      \foreach \x/\y/\lab in {
        -1.55/-0.70/{x_1},
        -0.55/0.85/{x_2},
         0.85/0.55/{x_3},
         1.45/-0.55/{x_4}
      }{
        \node[insertion] at (\x,\y) {};
        \node[violet!80!black,above right] at (\x,\y) {\(\lab\)};
      }
      \draw[blue!65!black,line width=0.7pt] (-1.55,-0.70)--(-0.55,0.85);
      \draw[blue!65!black,line width=0.7pt] (0.85,0.55)--(1.45,-0.55);
      \draw[orange!85!black,line width=0.7pt] (-1.55,-0.70)--(0.85,0.55);
      \draw[orange!85!black,line width=0.7pt] (-0.55,0.85)--(1.45,-0.55);
    \end{scope}

    \draw[arrow] (2.80,0)--(4.15,0);
    \node[align=center] at (3.47,0.52) {choose a\\conformal frame};

    \begin{scope}[shift={(6.80,0)}]
      \node[font=\small] at (0,2.0) {two-plane frame};
      \draw[thick] (-2.15,-1.15)--(2.15,-1.15);
      \draw[thick] (-1.65,-1.35)--(-1.65,1.35);
      \node[insertion] at (-1.65,-1.15) {};
      \node[violet!80!black,below] at (-1.65,-1.15) {\(0\)};
      \node[insertion] at (-0.55,0.45) {};
      \node[violet!80!black,above] at (-0.55,0.45) {\(z\)};
      \node[insertion] at (0.75,-1.15) {};
      \node[violet!80!black,below] at (0.75,-1.15) {\(1\)};
      \draw[arrow,violet!80!black] (1.25,-1.15)--(1.95,-1.15);
      \node[violet!80!black,below] at (1.95,-1.15) {\(\infty\)};
      \draw[blue!65!black,line width=0.7pt] (-1.65,-1.15)--(-0.55,0.45);
      \draw[orange!85!black,line width=0.7pt] (-1.65,-1.15)--(0.75,-1.15);
      \node[blue!65!black] at (-0.95,-0.25) {\(u=z\bar z\)};
      \node[orange!85!black] at (0.45,-0.55) {\(v=(1-z)(1-\bar z)\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{A scalar four-point function has two conformal invariants.  A
  conformal frame places the points at \(0,z,1,\infty\) in a two-plane.}
  \label{fig:four-point-conformal-frame}
\end{figure}

\begin{proposition}[Scalar four-point prefactor covariance]
\label{prop:scalar-four-point-prefactor-covariance}
Let \(\mathcal O_i\) be scalar primaries of dimensions \(\Delta_i\).  Every
separated conformally covariant four-point function can be written uniquely in
the form
\begin{align}
  &\left\langle
    \mathcal O_1(x_1)\mathcal O_2(x_2)
    \mathcal O_3(x_3)\mathcal O_4(x_4)
  \right\rangle
  \notag\\
  &\quad =
  \frac{\mathcal G(u,v)}
  {|x_{12}|^{\Delta_1+\Delta_2}
   |x_{34}|^{\Delta_3+\Delta_4}}
  \left(\frac{|x_{24}|}{|x_{14}|}\right)^{\Delta_1-\Delta_2}
  \left(\frac{|x_{14}|}{|x_{13}|}\right)^{\Delta_3-\Delta_4}.
  \label{eq:scalar-four-point-general}
\end{align}
For identical scalar primaries of dimension \(\Delta\), this simplifies to
\begin{equation}
  \left\langle
    \mathcal O(x_1)\mathcal O(x_2)
    \mathcal O(x_3)\mathcal O(x_4)
  \right\rangle
  =
  \frac{\mathcal G(u,v)}
       {|x_{12}|^{2\Delta}|x_{34}|^{2\Delta}}.
  \label{eq:identical-scalar-four-point}
\end{equation}
\end{proposition}

\begin{proof}
The quotient of a scalar four-point function by the displayed prefactor is
invariant under translations and rotations because it is built from
distances.  Under a uniform dilatation, the prefactor has degree
\(-(\Delta_1+\Delta_2+\Delta_3+\Delta_4)\), the required total degree.  It
remains to check the local weights under inversion.  Since
\[
  |x_{ij}|\longmapsto \frac{|x_{ij}|}{|x_i|\,|x_j|},
\]
write the prefactor as \(\prod_{i<j}|x_{ij}|^{a_{ij}}\), with
\[
\begin{gathered}
  a_{12}=-(\Delta_1+\Delta_2),
  \qquad
  a_{34}=-(\Delta_3+\Delta_4),
  \qquad
  a_{24}=\Delta_1-\Delta_2,\\
  a_{14}=-(\Delta_1-\Delta_2)+(\Delta_3-\Delta_4),
  \qquad
  a_{13}=-(\Delta_3-\Delta_4),
  \qquad
  a_{23}=0 .
\end{gathered}
\]
The inversion power at \(x_i\) is \(-\sum_{j\ne i}a_{ij}\).  Direct
substitution gives \(2\Delta_i\) for each \(i=1,\ldots,4\).  This is exactly
the inversion weight of four scalar primaries because
\(\Omega_I(x_i)=|x_i|^{-2}\).  Therefore the remaining factor is a
dimensionless conformal invariant and hence a function of \(u,v\) on the
generic configuration space by
Theorem~\ref{thm:four-point-cross-ratio-completeness}.  Uniqueness follows
because the prefactor is nonzero on \(\operatorname{Conf}_4(\R^D)\).  The
identical-scalar formula is the specialization
\(\Delta_1=\cdots=\Delta_4=\Delta\).
\end{proof}

The function \(\mathcal G(u,v)\) is dynamical CFT data.  The next stage is to
determine how the operator product expansion encodes local multiplication of
operators and how associativity constrains the allowed spectrum and OPE
coefficients.  Detailed computational machinery for solving crossing equations
is deferred from the core monograph until conformal theory is rebuilt as a
separate systematic development.
