\chapter{Analyticity, Crossing, and Landau Singularities}
\label{ch:volume-ii-analyticity-crossing-landau}
\ConventionsLorentzian

The analytic amplitudes in this chapter are boundary values of exact
Hilbert-space time-ordered or Wightman distributions when the stated
spectral, locality, and LSZ hypotheses are being used.  The Landau and
Feynman-parameter discussions are instead coefficientwise perturbative graph
analyses of regulated time-ordered kernels.  This is the standing convention
of Definition~\ref{def:scattering-time-ordered-correlator-status}.


\section{Physical Region and First Sheet}
\label{sec:physical-region-and-first-sheet}

Definition~\ref{def:stable-channel-sheet-convention} introduced the sheet
terminology at the first point where bound-state and resonance poles required
it.  This section develops the same convention for the general identical
massive \(2\to2\) amplitude and fixes the physical \(s\)-channel region used
in the analyticity discussion.

Consider \(2\to2\) scattering of identical stable scalar particles of mass
\(m\).  The metric is mostly plus, so a physical mass-\(m\) momentum \(k\)
obeys \(k^2=-m^2\).  Let the connected invariant amplitude be denoted by
\[
  \mathcal M(s,t),
\]
where
\[
  s=-(k_1+k_2)^2,\qquad
  t=-(k_1-k_3)^2,\qquad
  u=-(k_1-k_4)^2=4m^2-s-t .
\]
Here \(k_1,k_2\) are incoming and \(k_3,k_4\) are outgoing physical
momenta, in the convention of
Definition~\ref{def:mostly-plus-mandelstam-variables}.  Equivalently, all four
external momenta may be taken incoming after
replacing outgoing momenta by their negatives; the above definitions are the
physical scattering convention.  With \(k_i^2=-m^2\), the mostly-plus
Mandelstam identity is
\[
  s+t+u=4m^2 .
\]

In the center-of-mass frame,
\[
  s=E^2=4(k^2+m^2),
  \qquad
  t=-2k^2(1-\cos\theta),
  \qquad
  u=-2k^2(1+\cos\theta).
\]
Thus the physical \(s\)-channel region is
\[
  t\le0,\qquad
  u\le0,\qquad
  s\ge4m^2-t .
\]
The last inequality is the statement \(-t\le s-4m^2\).  For fixed
\(s\ge4m^2\), the allowed physical interval is
\[
  -(s-4m^2)\le t\le0 .
\]
For fixed real \(t<0\), the physical upper edge begins at
\[
  s=4m^2-t,
\]
because this is the value for which \(u=0\).  The analytic function whose
boundary value is the physical amplitude also has the lower \(s\)-channel
threshold at \(s=4m^2\); for \(4m^2<s<4m^2-t\) the point is on the same
right-hand cut but outside the physical \(s\)-channel scattering interval for
that fixed \(t\).  Figure~\ref{fig:fixed-t-physical-edge-cuts} fixes the
corresponding \(s\)-plane branch conventions.

\begin{figure}[!ht]
\centering
\begin{tikzpicture}[>=Stealth,scale=0.92]
  \begin{scope}
    \node[draw,circle,fill=gray!15,minimum size=0.78cm,inner sep=0pt] (a) at (0,0) {};
    \draw[->,line width=0.9pt] (-1.75,0.75)--(a.150);
    \draw[->,line width=0.9pt] (-1.75,-0.75)--(a.210);
    \draw[->,line width=0.9pt] (a.30)--(1.75,0.75);
    \draw[->,line width=0.9pt] (a.330)--(1.75,-0.75);
    \node at (0,0) {\(\mathcal M\)};
    \draw[dashed] (0,0)--(1.45,0);
    \draw[blue!70!black,->,line width=0.9pt]
      (0.85,0) arc[start angle=0,end angle=24,radius=0.85];
    \node[blue!70!black] at (1.05,0.3) {\(\theta\)};
    \node[align=center] at (0,-1.55)
      {\(s=4(k^2+m^2)\)\\
       \(t=-2k^2(1-\cos\theta)\)};
  \end{scope}
  \begin{scope}[xshift=6.5cm]
    \draw[->,line width=0.75pt] (-0.3,0)--(5.2,0) node[right] {\(s\)};
    \draw[->,line width=0.75pt] (0,-1.45)--(0,1.85)
      node[above] {\(\operatorname{Im}s\)};
    \draw[red!75!black,decorate,
      decoration={snake,amplitude=0.55mm,segment length=2.7mm},
      line width=1.05pt] (2.05,0)--(4.95,0);
    \node[red!75!black,below] at (2.05,-0.15) {\(4m^2\)};
    \draw[orange!85!black,decorate,
      decoration={snake,amplitude=0.55mm,segment length=2.7mm},
      line width=1.05pt] (0.95,0)--(-0.1,0);
    \node[orange!85!black,below] at (0.95,-0.15) {\(-t\)};
    \fill[blue!75!black] (1.45,0.28) circle (2.1pt);
    \fill[blue!75!black] (1.68,0.28) circle (2.1pt);
    \node[blue!75!black,align=center,font=\small] at (1.55,0.85)
      {possible\\bound poles};
    \draw[purple!80!black,->,line width=1.0pt]
      (2.25,0.62) .. controls (3.05,1.18) and (3.82,0.92) .. (4.18,0.15);
    \node[purple!80!black,align=center,font=\small] at (3.52,1.18)
      {physical\\upper edge};
  \end{scope}
\end{tikzpicture}
\caption{Fixed-\(t\) physical edge and cuts in the \(s\)-plane.  The left panel
fixes the center-of-mass kinematics and the angular variable.  The right panel
shows the first-sheet right-hand cut, the crossed cut, possible stable poles,
and the upper-edge boundary value used for the physical amplitude.}
\label{fig:fixed-t-physical-edge-cuts}
\end{figure}

For fixed real \(t<0\), the physical amplitude is the upper-edge boundary
value
\[
  \lim_{\epsilon\to0^+}\mathcal M(s+\ii\epsilon,t)
\]
on the first sheet.  In this monograph the terms \emph{first sheet} and
\emph{physical sheet} mean the same analytic branch: the germ reached from the
Euclidean or Feynman domain by the prescribed \(+\ii0\) continuation without
passing through any unitarity cut.  This is a statement about a branch of the
complexified invariant variables, not a synonym for the physical region of real
external momenta.  In the fixed-\(t\) \(s\)-plane this branch has the right-hand
\(s\)-channel cut
\[
  C_s=[4m^2,\infty),
\]
and, because \(u=4m^2-s-t\), the crossed \(u\)-channel cut
\[
  C_u=(-\infty,-t] .
\]
A stable state of mass \(M_B<2m\) coupled to the two external scalars gives a
first-sheet pole at \(s=M_B^2\) in the corresponding channel.  In the crossed
channel this same statement appears as a pole at
\[
  s=4m^2-t-M_B^2 .
\]
If the stable state is composite, this pole is the same spectral pole whose
four-point residue is factorized in
Theorem~\ref{thm:bethe-salpeter-pole-factorization}.  In a two-constituent
description its mass is equivalently determined by the homogeneous
Bethe--Salpeter condition
\eqref{eq:homogeneous-bethe-salpeter-equation}, or by
\((\mathbf 1-\widehat V_P)\Psi_B=0\) at \(P^2=-M_B^2\).  Crossing changes the
Mandelstam coordinate in which the pole is seen; it does not change the Hilbert
space state or the residue matrix element from which the pole term is built.
For the purposes used here, the first-sheet data consist of this cut plane
together with the possible pole data and with any additional first-sheet
singularities allowed by the full spectral problem.

A \emph{second sheet} is always defined relative to a specified cut.  It is the
adjacent analytic germ obtained by continuing the first-sheet germ through that
cut.  Near a single two-particle threshold, with
\[
  \beta(s)=\sqrt{1-\frac{s_{\mathrm{th}}}{s}}
\]
chosen so that \(\beta(s)>0\) on the physical upper edge \(s>s_{\mathrm{th}}\),
the continuation through the corresponding unitarity cut sends
\(\beta\mapsto-\beta\).  With several thresholds there is no cut-independent
meaning of ``the'' second sheet; the sheet is labelled by the set of cuts
through which the analytic germ has been continued, or equivalently by the
monodromy choices for the associated channel momenta.

\section{Real Analyticity and Crossing}

For the analytic discussion in this chapter assume a mass gap, Poincare
invariance, locality, unitarity, and the existence of stable one-particle
external states to which LSZ applies.  These assumptions specify the
structural input used here, while the maximal analytic domain remains an
additional problem.  The continuation used in this chapter is an on-shell
continuation in the invariant variables.  The external legs remain on the
complexified mass shell; the variables \(s,t,u\) are complex coordinates on
the algebraic surface
\[
  \mathfrak S_m
  =
  \{(s,t,u)\in\C^3:\ s+t+u=4m^2\},
\]
with \(u=4m^2-s-t\).  The starting object on a real physical edge is the
connected LSZ amplitude as a tempered distribution in wave-packet momenta,
after the overall momentum-conserving delta function and external pole
residues have been extracted.  Away from thresholds, stable-particle poles,
and other singular loci, this distribution is represented by boundary values
of holomorphic functions on tube-like domains in \(\mathfrak S_m\).

\begin{definition}[Tempered holomorphic amplitude germ]
\label{def:tempered-holomorphic-amplitude-germ}
Let \(\mathcal D\subset\mathfrak S_m\) be a connected open set and let
\(\mathcal D_\R\) be a real edge of \(\mathcal D\).  A holomorphic amplitude
germ on \(\mathcal D\) is an element
\[
  F\in\mathcal O_{\mathrm{poly}}(\mathcal D)
\]
with the following local polynomial-growth property: for every relatively
compact \(K\Subset\mathcal D\) that approaches \(\mathcal D_\R\) inside a
fixed open cone of imaginary directions, there are constants \(C_K,N_K\) such
that
\[
  |F(z)|\le C_K(1+|z|)^{N_K}
  \operatorname{dist}(z,\partial\mathcal D)^{-N_K}.
\]
Its boundary value on an open real set \(U\subset\mathcal D_\R\) is the
distribution
\[
  \operatorname{bv}F\in\mathcal D'(U)
\]
defined by
\[
  \langle \operatorname{bv}F,\varphi\rangle
  =
  \lim_{\lambda\downarrow0}
  \int_U F(x+\ii\lambda\eta)\varphi(x)\,\dd x,
  \qquad \varphi\in C_c^\infty(U),
\]
whenever the limit exists and is independent of \(\eta\) inside the chosen
cone.  On noncompact real edges the same polynomial-growth condition permits
Schwartz test functions and gives a tempered boundary value in
\(\mathcal S'\) in the chosen real coordinate chart.  Equality of two physical
amplitudes on \(U\) means equality of these boundary values in the strong dual
topology of \(\mathcal D'(U)\), or of \(\mathcal S'\) when the noncompact
tempered version is used.
\end{definition}

\subsection{Primitive Tube Domains from Wightman Data}
\label{subsec:primitive-wightman-tube-domains}

We now isolate the first domain theorem used in scattering analyticity.  It is
not a perturbative statement.  It is the theorem that converts the spectral
condition and locality of Wightman fields into holomorphic domains for
correlation functions.  The later Lehmann--Martin and Jin--Martin arguments
start from this primitive domain and add the further retarded-commutator,
partial-wave, and unitarity input needed for fixed-\(t\) dispersion theory.

Let \(\Phi_a(x)\), \(a\in A\), be scalar Wightman fields for simplicity of
notation.  Spin and finite-dimensional internal labels only insert the
corresponding covariance matrices and do not change the domain argument.  For
an ordering \(\sigma\in S_n\), define the ordered \(n\)-point distribution
\[
  W_\sigma(x_1,\ldots,x_n)
  =
  \langle\Omega,
    \Phi_{\sigma(1)}(x_{\sigma(1)})
    \cdots
    \Phi_{\sigma(n)}(x_{\sigma(n)})
  \Omega\rangle .
\]
Translation invariance allows one to remove the center-of-mass variable.  For
the displayed ordering, put
\[
  \xi_j=x_{\sigma(j)}-x_{\sigma(j+1)},
  \qquad j=1,\ldots,n-1,
\]
and write the reduced distribution as \(w_\sigma(\xi_1,\ldots,\xi_{n-1})\).
Let \(P^\mu\) be the self-adjoint generators of spacetime translations and
let \(\overline V_+\) denote the closed future light cone in momentum space,
with the mostly-plus mass convention fixed in Volume I.

\begin{theorem}[Primitive Wightman tube-domain theorem]
\label{thm:primitive-wightman-tube-domain}
Assume the Wightman hypotheses for the fields appearing above: a
translation-invariant vacuum, strong continuity of translations, the spectrum
condition
\[
  \operatorname{Spec}(P)\subset \overline V_+,
\]
and temperedness of all vacuum matrix elements.  Then, for every ordering
\(\sigma\),
\begin{equation}
  \operatorname{supp}\widehat w_\sigma
  \subset
  (\overline V_+)^{n-1}
  \label{eq:ordered-wightman-spectral-support}
\end{equation}
in the reduced Fourier variables conjugate to
\((\xi_1,\ldots,\xi_{n-1})\).  Consequently \(w_\sigma\) is the tempered
boundary value of a holomorphic function
\[
  \mathcal W_\sigma(z_1,\ldots,z_{n-1}),
  \qquad
  z_j=\xi_j+\ii\eta_j ,
\]
on the tube
\begin{equation}
  T_{n-1}
  =
  \{(\xi_j+\ii\eta_j)_{j=1}^{n-1}\mid
    \eta_j\in V_+^\circ,\ j=1,\ldots,n-1\},
  \label{eq:primitive-forward-tube}
\end{equation}
with the same Fourier--Laplace sign convention as
Proposition~\ref{prop:fourier-laplace-tempered-boundary-values}.  More
precisely, for every closed subcone \(K\Subset(V_+^\circ)^{n-1}\) the tested
functions
\[
  f\longmapsto
  \int \mathcal W_\sigma(\xi+\ii\eta)f(\xi)\,\dd^{D(n-1)}\xi
\]
are bounded by finitely many Schwartz seminorms of \(f\), with constants
polynomial in the reciprocal distance of \(\eta\) from the boundary of the
cone, and converge in \(\mathcal S'\) to \(w_\sigma\) as
\(\eta\to0\) inside \(K\).
\end{theorem}

\begin{proof}
We first prove the support statement.  It is enough to test
\(\widehat w_\sigma\) against a compactly supported smooth function
\(\widehat f(p_1,\ldots,p_{n-1})\) whose support is disjoint from
\((\overline V_+)^{n-1}\).  Choose Schwartz functions
\[
  f(\xi_1,\ldots,\xi_{n-1})
  =
  \int
  \ee^{-\ii\sum_j p_j\cdot \xi_j}
  \widehat f(p_1,\ldots,p_{n-1})
  \prod_j\frac{\dd^Dp_j}{(2\pi)^D}.
\]
The smeared reduced Wightman distribution is obtained by inserting the
joint spectral resolution of the translation representation between adjacent
fields.  On the common invariant domain this gives the weak identity
\begin{equation}
\begin{aligned}
  \langle \widehat w_\sigma,\widehat f\rangle
  &=
  \int
  \widehat f(p_1,\ldots,p_{n-1})\,
  \dd\mu_\sigma(p_1,\ldots,p_{n-1}),
\end{aligned}
  \label{eq:wightman-spectral-measure-pairing}
\end{equation}
where the scalar distributional measure \(\mu_\sigma\) is built from the
successive spectral projections of the translation representation applied to
the vectors generated by the ordered fields.  The \(j\)-th momentum \(p_j\)
is the total momentum carried by the state between the \(j\)-th and
\((j+1)\)-st inserted fields.  By the spectrum condition each such
intermediate momentum lies in \(\overline V_+\).  Therefore the right hand
side of \eqref{eq:wightman-spectral-measure-pairing} vanishes whenever
\(\operatorname{supp}\widehat f\cap(\overline V_+)^{n-1}=\varnothing\),
which is exactly
\eqref{eq:ordered-wightman-spectral-support}.

For completeness we spell out why the preceding spectral-resolution formula
is legitimate for distributional fields.  First smear each field with a
Schwartz function and use the spectral theorem for the strongly continuous
unitary translation group between the smeared field operators.  The
translation covariance relation
\[
  U(a)\Phi_b(h)U(a)^{-1}=\Phi_b(h_a),
  \qquad h_a(x)=h(x-a),
\]
turns translations of adjacent insertions into the Fourier factors in
\eqref{eq:wightman-spectral-measure-pairing}.  The Wightman temperedness
hypothesis gives continuity in the Schwartz topology, so the identity extends
from tensor-product test functions to all Schwartz tests in the reduced
variables and then by density to the displayed \(\widehat f\).

The holomorphic continuation is now the Fourier--Laplace theorem for
tempered distributions carried by a closed convex cone, applied to the cone
\((\overline V_+)^{n-1}\).  The exponential damping in the imaginary
directions \(\eta_j\in V_+^\circ\) makes the Fourier transform into a
holomorphic function of \(z_j=\xi_j+\ii\eta_j\).  The finite-order
Schwartz-seminorm estimate for \(\widehat w_\sigma\) gives the displayed
polynomial tube bounds after differentiating the exponential multiplier.  The
same estimate, with the exponential multiplier tending to \(1\) in the
Schwartz topology after testing, gives the boundary value
\(\mathcal W_\sigma(\cdot+\ii\eta)\to w_\sigma\) in \(\mathcal S'\).  This
is precisely the construction written abstractly in
Proposition~\ref{prop:fourier-laplace-tempered-boundary-values}.
\end{proof}

The primitive tube is only the first part of the domain.  Locality determines
how different ordered tubes are glued.  The relevant real edge is the Jost
set.  In the reduced variables of an ordering, define
\[
  C(\xi)
  =
  \left\{
    \sum_{j=1}^{n-1}\lambda_j\xi_j\ \middle|
    \lambda_j\ge0
  \right\}.
\]
A real reduced configuration \(\xi\) is a Jost configuration if every
nonzero vector in \(C(\xi)\) is spacelike.  With the mostly-plus convention,
this means \(v^2>0\) for all \(0\ne v\in C(\xi)\).  In particular, every
difference \(x_{\sigma(i)}-x_{\sigma(k)}=\xi_i+\cdots+\xi_{k-1}\) between
two ordered insertion points is spacelike.

\begin{theorem}[Jost-edge gluing of primitive tubes]
\label{thm:jost-edge-gluing-primitive-tubes}
Assume, in addition to the hypotheses of
Theorem~\ref{thm:primitive-wightman-tube-domain}, graded local commutativity
on the common Wightman domain:
\[
  \Phi_a(f)\Phi_b(g)
  =
  (-1)^{|a||b|}\Phi_b(g)\Phi_a(f)
\]
whenever the supports of \(f\) and \(g\) are spacelike separated.  Let
\(\sigma'\) be obtained from \(\sigma\) by one adjacent transposition of
field labels whose two insertion points are spacelike throughout an open Jost
neighbourhood \(U_J\).  Then the two boundary distributions
\(\operatorname{bv}\mathcal W_\sigma\) and
\(\operatorname{bv}\mathcal W_{\sigma'}\), expressed in the same real
coordinates on \(U_J\), agree there up to the Koszul sign of the
transposition.  Hence the distributional edge-of-the-wedge theorem gives a
single holomorphic function on the envelope of holomorphy of the union of the
two adjacent primitive tubes and the common Jost edge.  Iterating adjacent
transpositions gives a well-defined holomorphic germ on the primitive
Bros--Epstein--Glaser envelope generated by all ordered Wightman tubes and
their Jost-edge identifications.
\end{theorem}

\begin{proof}
Choose test functions supported in a coordinate neighbourhood of \(U_J\).
For the adjacent pair being transposed, every point in the product support
has spacelike separation.  Graded local commutativity therefore permits the
two smeared fields to be exchanged on the common invariant domain, with the
factor \((-1)^{|a||b|}\).  Leaving all other insertions fixed gives equality
of the two ordered Wightman distributions on \(U_J\), again with the same
Koszul factor.  This proves equality of the real boundary values on the Jost
edge.

By Theorem~\ref{thm:primitive-wightman-tube-domain}, each ordered
distribution is the boundary value of a holomorphic function with polynomial
tube bounds.  The distributional edge-of-the-wedge theorem applies exactly
to such a pair: holomorphic functions in two opposite or adjacent wedges with
equal distributional boundary values on a common real edge extend to the
envelope of holomorphy of their union.  The extension is unique because a
holomorphic function with zero distributional boundary value on an open edge
vanishes in the connected envelope.  Applying this step to a sequence of
adjacent transpositions gives the asserted primitive envelope.  If two
different transposition sequences relate the same orderings, uniqueness on
the overlaps makes the resulting germs identical; thus the iterated
continuation is path independent inside the generated envelope.
\end{proof}

\begin{remark}[Output of the primitive domain theorem for four-point scattering]
\label{rem:primitive-domain-output-for-scattering}
For a massive Wightman theory with stable scalar external particles, the
connected four-point Wightman distributions possess a crossing-symmetric
primitive analytic germ obtained from Theorems
\ref{thm:primitive-wightman-tube-domain} and
\ref{thm:jost-edge-gluing-primitive-tubes}.  After separating the connected
part and the external one-particle poles, LSZ boundary values in the physical
\(s\)-, \(t\)-, and \(u\)-channel regions are boundary values of this germ on
the corresponding real edges whenever those LSZ limits exist.
Indeed, for \(n=4\), the preceding two theorems apply to each ordering and
glue the ordered tubes along Jost edges.  Passing to the connected part removes
finite sums of products of lower-point distributions with the same spectral
support and locality properties.  The statement about scattering is not an
additional analyticity theorem: it is the conditional transfer of this
connected four-point germ through the already defined LSZ boundary operation,
with wave-packet smearing, removal of the external one-particle pole factors,
and the mass-shell limit taken only on edges where that limit exists.
\end{remark}

\subsection{Causal Commutators and the Dyson Representation}
\label{subsec:causal-commutator-dyson-representation}

The next domain theorem uses the commutator rather than the ordered
Wightman functions.  It is the analytic reason that a coincidence region in
momentum space can be converted into a representation by mass hyperboloids.
This is the Jost--Lehmann--Dyson mechanism in the form needed for scattering
amplitudes.

Let \(C(x)\in\mathcal S'(\mathbb M^4)\) be a scalar tempered distribution
whose support is contained in the closed causal cone
\[
  \overline V_{\rm caus}
  =
  \{x\in\mathbb M^4\mid x^2\le0\}
\]
in the mostly-plus convention.  In QFT applications
\begin{equation}
  C(x)
  =
  \langle Q_f,
    [J_1(x/2),J_2(-x/2)]
  Q_i\rangle
  \label{eq:causal-commutator-matrix-coefficient}
\end{equation}
with \(J_i\) local current operators and with \(Q_i,Q_f\) finite-energy
vectors.  Local commutativity gives the support condition.  Let
\[
  \widehat C(q)=\langle C(x),\ee^{\ii q\cdot x}\rangle
\]
be its Fourier transform, again as a tempered distribution.

The representation also uses a momentum-space coincidence region.  Let
\begin{equation}
  R=\{q\in\mathbb R^4\mid h_-(\vec q)<q^0<h_+(\vec q)\}
  \label{eq:dyson-coincidence-slab}
\end{equation}
where \(h_\pm:\mathbb R^3\to\mathbb R\) are continuous functions satisfying
\[
  |h_\pm(\vec q)-h_\pm(\vec q\,')|
  \le |\vec q-\vec q\,'|
\]
and \(h_-(\vec q)<h_+(\vec q)\).  Thus the boundary graphs are spacelike or
null.  Assume
\begin{equation}
  \widehat C(q)=0
  \qquad q\in R
  \label{eq:causal-commutator-vanishing-region}
\end{equation}
as a distributional restriction to \(R\).  For \(u\in\mathbb R^4\) and
\(\mu\ge0\), define the two mass-hyperboloid sheets
\[
  H_\pm(u,\mu)
  =
  \{q\in\mathbb R^4\mid (q-u)^2=-\mu^2,\ \pm(q^0-u^0)>0\}.
\]
The pair \((u,\mu)\) is called \(R\)-admissible if
\begin{equation}
  q\in H_+(u,\mu)\Rightarrow q^0\ge h_+(\vec q),
  \qquad
  q\in H_-(u,\mu)\Rightarrow q^0\le h_-(\vec q).
  \label{eq:dyson-admissible-hyperboloid}
\end{equation}
Let \(S_R\subset\mathbb R^4\times[0,\infty)\) be the closed set of
\(R\)-admissible pairs.

\begin{proposition}[Distributional Cauchy formula for the six-dimensional wave equation]
\label{prop:six-dimensional-distributional-cauchy-formula}
Let \(Y=(\tau,\mathbf y)\in\mathbb R^{1,5}\), with
\(\mathbf y\in\mathbb R^5\), and let
\[
  \Box_{1,5}=\partial_\tau^2-\Delta_{\mathbf y}.
\]
Let \(F\in\mathcal S'(\mathbb R^{1,5})\) solve
\[
  \Box_{1,5}F=0
\]
in the distributional sense.  Fix an affine spacelike hyperplane
\(\Sigma_{\tau_*}=\{\tau=\tau_*\}\), and suppose the Cauchy traces
\[
  f=F|_{\Sigma_{\tau_*}},
  \qquad
  g=\partial_\tau F|_{\Sigma_{\tau_*}}
\]
exist as tempered distributions on \(\mathbb R^5\).  Let \(D_6\) be the odd
fundamental solution of the six-dimensional wave equation, normalized by
\begin{equation}
  \mathcal F_{\mathbf y}D_6(t,\cdot)(\mathbf k)
  =
  \frac{\sin(|\mathbf k|t)}{|\mathbf k|},
  \qquad
  \partial_tD_6(0,\mathbf y)=\delta(\mathbf y),
  \qquad
  D_6(0,\mathbf y)=0 .
  \label{eq:six-dimensional-wave-commutator-kernel}
\end{equation}
Then, as a tempered distribution in \(Y_0=(\tau_0,\mathbf y_0)\),
\begin{equation}
  F(Y_0)
  =
  \left\langle
    f(\mathbf y),
    \partial_{\tau}D_6(\tau_*-\tau_0,\mathbf y-\mathbf y_0)
  \right\rangle_{\mathbf y}
  -
  \left\langle
    g(\mathbf y),
    D_6(\tau_*-\tau_0,\mathbf y-\mathbf y_0)
  \right\rangle_{\mathbf y}.
  \label{eq:six-dimensional-distributional-cauchy-formula}
\end{equation}
Equivalently, for an affine spacelike hyperplane
\(\Sigma\subset\mathbb R^{1,5}\) with future unit normal \(n\),
\begin{equation}
  F(Y_0)
  =
  \int_\Sigma
  \left(
    F(Y)\,\partial_{n_Y}D_6(Y-Y_0)
    -
    D_6(Y-Y_0)\,\partial_{n_Y}F(Y)
  \right)\dd\Sigma(Y),
  \label{eq:six-dimensional-geometric-cauchy-formula}
\end{equation}
where the displayed integral denotes the distributional pairing of the
Cauchy traces with the kernel.
\end{proposition}

\begin{proof}
It is enough to prove the formula for \(\tau_*=0\); translating \(\tau\)
gives the general case.  Taking the Fourier transform in the five spatial
variables converts the wave equation into
\begin{equation}
  \partial_\tau^2\widehat F(\tau,\mathbf k)
  +
  |\mathbf k|^2\widehat F(\tau,\mathbf k)=0
  \label{eq:six-dimensional-wave-spatial-fourier-ode}
\end{equation}
in \(\mathcal S'(\mathbb R_\tau\times\mathbb R^5_{\mathbf k})\).  A
distributional solution of this constant-coefficient ordinary differential
equation is uniquely determined by
\[
  \widehat F(0,\mathbf k)=\widehat f(\mathbf k),
  \qquad
  \partial_\tau\widehat F(0,\mathbf k)=\widehat g(\mathbf k),
\]
because the difference of two such solutions has zero Cauchy data and
therefore pairs to zero with every compactly supported test function after
solving the adjoint final-value problem.  The solution is
\begin{equation}
  \widehat F(\tau,\mathbf k)
  =
  \cos(|\mathbf k|\tau)\widehat f(\mathbf k)
  +
  \frac{\sin(|\mathbf k|\tau)}{|\mathbf k|}\widehat g(\mathbf k),
  \label{eq:six-dimensional-wave-fourier-solution}
\end{equation}
with the quotient interpreted by its smooth value \(\tau\) at
\(\mathbf k=0\).  Inverse Fourier transformation and the normalization
\eqref{eq:six-dimensional-wave-commutator-kernel} give
\eqref{eq:six-dimensional-distributional-cauchy-formula}.  The geometric
form follows by applying a real Lorentz transformation carrying the chosen
affine spacelike hyperplane to a constant-\(\tau\) hyperplane; the wave
operator, \(D_6\), the normal derivative, and the distributional pairing are
invariant under that transformation.
\end{proof}

\begin{proposition}[Auxiliary pushforward and the \(p=0\) momentum trace]
\label{prop:auxiliary-pushforward-momentum-trace}
Let
\[
  U\in\mathcal S'(\mathbb R_x^4\times\mathbb R_y^2)
\]
and define its Fourier transform by duality from the convention
\begin{equation}
  \Psi^\vee(x,y)
  =
  \int_{\mathbb R^4\times\mathbb R^2}
  \Psi(q,p)\ee^{\ii q\cdot x+\ii p\cdot y}
  \,\dd^4q\,\dd^2p,
  \qquad
  \langle\widehat U,\Psi\rangle=\langle U,\Psi^\vee\rangle .
  \label{eq:auxiliary-trace-fourier-convention}
\end{equation}
Choose \(\rho\in\mathcal S(\mathbb R^2_p)\) with
\(\int\rho(p)\,\dd^2p=1\), set
\begin{equation}
  \rho_\epsilon(p)=\epsilon^{-2}\rho(p/\epsilon),
  \qquad
  m_\epsilon(y)=\int_{\mathbb R^2}\rho_\epsilon(p)
  \ee^{\ii p\cdot y}\,\dd^2p .
  \label{eq:auxiliary-trace-mollifier}
\end{equation}
Suppose that for every \(\varphi\in\mathcal S(\mathbb R_x^4)\) the limit
\begin{equation}
  \langle\pi_{y*}U,\varphi\rangle
  :=
  \lim_{\epsilon\downarrow0}
  \langle U,\varphi(x)m_\epsilon(y)\rangle
  \label{eq:auxiliary-pushforward-definition}
\end{equation}
exists and is independent of the normalized approximate identity \(\rho\).
Then \(\pi_{y*}U\in\mathcal S'(\mathbb R_x^4)\), and the trace of
\(\widehat U\) at \(p=0\), defined by
\begin{equation}
  \left\langle \iota^*\widehat U,\psi\right\rangle
  :=
  \lim_{\epsilon\downarrow0}
  \left\langle
    \widehat U(q,p),
    \psi(q)\rho_\epsilon(p)
  \right\rangle ,
  \qquad
  \iota(q)=(q,0),
  \label{eq:auxiliary-momentum-trace-definition}
\end{equation}
exists for every \(\psi\in\mathcal S(\mathbb R_q^4)\) and obeys
\begin{equation}
  \iota^*\widehat U=\widehat{\pi_{y*}U}.
  \label{eq:auxiliary-pushforward-trace-identity}
\end{equation}
Conversely, if the limits in
\eqref{eq:auxiliary-momentum-trace-definition} exist for every
\(\psi\) and are continuous in the Schwartz topology, then
\eqref{eq:auxiliary-pushforward-definition} defines a tempered
\(y\)-pushforward and
\eqref{eq:auxiliary-pushforward-trace-identity} holds.
\end{proposition}

\begin{proof}
For \(\psi\in\mathcal S(\mathbb R_q^4)\), define
\begin{equation}
  \psi^\vee(x)=\int_{\mathbb R^4}\psi(q)\ee^{\ii q\cdot x}\,\dd^4q .
  \label{eq:auxiliary-trace-test-inverse}
\end{equation}
With the convention
\eqref{eq:auxiliary-trace-fourier-convention}, the definition of the
Fourier transform of a tempered distribution gives
\begin{equation}
\begin{aligned}
  \left\langle
    \widehat U(q,p),
    \psi(q)\rho_\epsilon(p)
  \right\rangle
  &=
  \left\langle
    U(x,y),
    \psi^\vee(x)m_\epsilon(y)
  \right\rangle .
\end{aligned}
  \label{eq:auxiliary-trace-pairing-identity}
\end{equation}
If the pushforward limit exists, the right side tends to
\[
  \langle\pi_{y*}U,\psi^\vee\rangle
  =
  \langle\widehat{\pi_{y*}U},\psi\rangle,
\]
which proves existence of the trace and
\eqref{eq:auxiliary-pushforward-trace-identity}.  Continuity follows because
the pointwise limit of the continuous functionals
\(\varphi\mapsto\langle U,\varphi(x)m_\epsilon(y)\rangle\) on the
Schwartz space is continuous by the uniform boundedness principle for
Frechet spaces; the Fourier transform is a topological automorphism of the
Schwartz space.

Conversely, the map
\(\psi\mapsto\psi^\vee\) is a topological automorphism of
\(\mathcal S(\mathbb R^4)\).  Given
\(\varphi\in\mathcal S(\mathbb R^4)\), choose
\(\psi\) with \(\psi^\vee=\varphi\).  If the limits in
\eqref{eq:auxiliary-momentum-trace-definition} exist continuously, then
\eqref{eq:auxiliary-trace-pairing-identity} defines the limit in
\eqref{eq:auxiliary-pushforward-definition}; the resulting functional on
\(\varphi\) is tempered, and its Fourier transform is precisely
\(\iota^*\widehat U\).  This proves the converse and the identity.
\end{proof}

\begin{definition}[Dyson-regular causal commutator]
\label{def:dyson-regular-causal-commutator}
A causal commutator distribution \(C\in\mathcal S'(\mathbb M^4)\) is
Dyson-regular if its Fourier transform is the restriction to the plane
\(p_1=p_2=0\) of a rotationally invariant distribution
\[
  F(q,p_1,p_2)\in\mathcal S'(\mathbb R^{1,5})
\]
with the following properties:
\begin{enumerate}
  \item \(F\) solves the six-dimensional wave equation
  \[
    \Box_{1,5}F=0
  \]
  in the distributional sense.
  \item On every affine spacelike hyperplane used in the Dyson argument, the
  traces \(F|_\Sigma\) and \(\partial_nF|_\Sigma\) exist as tempered
  distributions, and the distributional Cauchy formula
  \eqref{eq:six-dimensional-geometric-cauchy-formula} holds.
  \item If \(U\) is the inverse Fourier transform of \(F\), then
  \(U\) is supported on the six-dimensional light cone, the auxiliary
  \(y\)-pushforward \(\pi_{y*}U\) exists in the sense of
  Proposition~\ref{prop:auxiliary-pushforward-momentum-trace}, and
  \[
    \pi_{y*}U=\kappa C
  \]
  for a fixed nonzero normalization \(\kappa\).  Equivalently, after
  absorbing \(\kappa\) into the definition of the lift,
  \[
    \iota^*F=\widehat C,
    \qquad
    \iota(q)=(q,0,0).
  \]
\end{enumerate}
For a sufficiently regular function \(C(x)\) with causal support this lift is
represented by Dyson's formula
\[
  \widetilde C(x,y)=4\pi C(x)\delta(x_D^2-|y|^2),
  \qquad y\in\mathbb R^2,
\]
in the temporary mostly-minus notation \(x_D^2=(x^0)^2-|\vec x|^2\).  This
display is not a definition for an arbitrary Wightman distribution: multiplying
an arbitrary distribution by the light-cone delta is a microlocal operation and
requires a transversality or regularization argument.  The definition records
exactly the regularity datum needed for the Dyson proof; verifying it for a
specified class of unsmeared QFT matrix coefficients is a separate theorem,
not a notational convention.
\end{definition}

\begin{definition}[Support-pruned Dyson data in a coincidence slab]
\label{def:support-pruned-dyson-data}
Let \(R\) be the coincidence slab in
\eqref{eq:dyson-coincidence-slab}, and let \(S_R\) be the corresponding
closed set of admissible hyperboloids.  For a distribution
\[
  \Psi\in\mathcal D'(\mathbb R^4_u\times[0,\infty)_\mu)
\]
for which the pairing is defined, write
\begin{equation}
  (\mathcal H\Psi)(q)
  =
  \left\langle
    \Psi(u,\mu),
    \varepsilon(q^0-u^0)
    \delta\!\bigl((q-u)^2+\mu^2\bigr)
  \right\rangle_{u,\mu}
  \label{eq:dyson-hyperboloid-transform}
\end{equation}
for the Dyson hyperboloid transform.  A Dyson-regular commutator has
support-pruned Dyson data relative to \(R\) if the Cauchy-data distribution
\(\Psi_\Sigma\) produced from its six-dimensional wave lift can be replaced,
modulo the kernel of \(\mathcal H\), by a distribution \(\Psi_R\) satisfying
\begin{equation}
  \operatorname{supp}\Psi_R\subset S_R,
  \qquad
  \mathcal H\Psi_R=\mathcal H\Psi_\Sigma .
  \label{eq:support-pruned-dyson-data}
\end{equation}
The condition is independent of the chosen affine Cauchy hyperplane in the
following sense: changing the hyperplane changes \(\Psi_\Sigma\) by a
distribution in the kernel of \(\mathcal H\), because both Cauchy surfaces
reconstruct the same six-dimensional wave solution by
Proposition~\ref{prop:six-dimensional-distributional-cauchy-formula}.
\end{definition}

\begin{proposition}[Axis trace nonuniqueness]
\label{prop:dyson-axis-trace-nonuniqueness}
Let the six-dimensional wave operator be written in affine coordinates
\[
  Y=(\tau,\xi,p)\in\mathbb R_\tau\times\mathbb R^3_\xi\times\mathbb R^2_p,
  \qquad
  \Box_{1,5}=\partial_\tau^2-\Delta_\xi-\Delta_p .
\]
There exists a nonzero \(O(2)\)-invariant tempered smooth solution
\[
  F\in C^\infty(\mathbb R^{1,5})
\]
of \(\Box_{1,5}F=0\) such that
\begin{equation}
  F(\tau,\xi,0)=0
  \qquad
  \text{for } |\tau|<1,\ \xi\in\mathbb R^3 .
  \label{eq:axis-trace-vanishing-example}
\end{equation}
Thus vanishing of the auxiliary-momentum trace on a coincidence slab does not
by itself imply that the six-dimensional Cauchy data vanish in the region of
Cauchy-data space whose hyperboloids meet that slab.
\end{proposition}

\begin{proof}
Choose a nonzero function
\[
  a\in C_c^\infty((2,3))
\]
and define radial Cauchy data on the auxiliary \(p\)-plane by
\[
  f(p)=a(|p|),
  \qquad
  g(p)=0 .
\]
Let \(F_0(\tau,p)\) be the solution of the \(1+2\)-dimensional wave equation
\[
  (\partial_\tau^2-\Delta_p)F_0=0,
  \qquad
  F_0(0,p)=f(p),\qquad
  \partial_\tau F_0(0,p)=g(p).
\]
The Cauchy problem with smooth compactly supported data has a unique smooth
solution, obtained for instance by Fourier transform in \(p\).  Rotational
invariance of the operator and uniqueness imply that \(F_0(\tau,p)\) is
\(O(2)\)-invariant.  The finite-propagation estimate follows from the local
energy identity.  For a solution \(u\) of
\((\partial_\tau^2-\Delta_p)u=0\), the stress tensor
\[
  e=\frac12\bigl((\partial_\tau u)^2+|\nabla_pu|^2\bigr),
  \qquad
  j=-\partial_\tau u\,\nabla_pu
\]
satisfies \(\partial_\tau e+\nabla_p\cdot j=0\).  Integrating this identity
over a truncated backward cone with vertex \((\tau_0,p_0)\) gives
\[
  E(s_2)+\int_{\text{null side}}\frac12
  \bigl(\partial_\tau u+\hat n\cdot\nabla_pu\bigr)^2\,\dd\sigma
  =
  E(s_1),
\]
where \(E(s)\) is the energy on the disk
\(|p-p_0|\le \tau_0-s\) at time \(s\) and \(\hat n\) is the outward radial
unit vector on the side of the cone.  The side flux is nonnegative, so zero
Cauchy data on the base disk force \(u\) to vanish in the cone.  Hence the
value at \((\tau,p_0)\) depends only on the data in
\[
  \{p\in\mathbb R^2: |p-p_0|\le |\tau|\}.
\]
For \(p_0=0\) and \(|\tau|<1\), this disk is disjoint from
\(\operatorname{supp}f\subset\{|p|>2\}\) and from
\(\operatorname{supp}g\).  Hence \(F_0(\tau,0)=0\) for \(|\tau|<1\).
On the other hand \(F_0(0,p)=a(|p|)\) is not identically zero.

Define
\[
  F(\tau,\xi,p)=F_0(\tau,p).
\]
Then \(F\) is independent of \(\xi\), solves
\(\partial_\tau^2F-\Delta_\xi F-\Delta_pF=0\), is \(O(2)\)-invariant in
the auxiliary variables, and satisfies
\eqref{eq:axis-trace-vanishing-example}.  Since \(F_0\) is smooth and has at
most polynomial growth as a distribution under the Fourier-transform solution
formula, its pullback constant in \(\xi\) is tempered on
\(\mathbb R^{1,5}\).  This proves the claim.
\end{proof}

\begin{definition}[Dyson regularity modulo contact terms]
\label{def:dyson-regular-modulo-contact}
A causal commutator \(C\in\mathcal S'(\mathbb M^4)\) is Dyson-regular
modulo contact terms if there is a finite-order distribution \(L\) supported
at \(x=0\) such that \(C-L\) is Dyson-regular in the sense of
Definition~\ref{def:dyson-regular-causal-commutator}.  Equivalently,
\(\widehat C\) has the Dyson hyperboloid representation of
Theorem~\ref{thm:dyson-causal-commutator-representation} plus a polynomial in
\(q\).  This is the form actually used after source-current reduction,
because applying differential wave operators to time-ordered products can
produce local terms at coincident points.
\end{definition}

\begin{definition}[Local scaling degree]
\label{def:local-scaling-degree}
Let \(T\in\mathcal D'(\mathbb R^n\setminus\{0\})\).  For
\(\varphi\in C_c^\infty(\mathbb R^n)\), set
\[
  \varphi_\lambda(z)=\lambda^{-n}\varphi(z/\lambda),
  \qquad 0<\lambda<1 .
\]
The scaling degree of \(T\) at the origin is
\begin{equation}
  \operatorname{sd}_0(T)
  =
  \inf\left\{
    \omega\in\mathbb R\ \middle|\
    \lambda^{\omega+\epsilon}\langle T,\varphi_\lambda\rangle\to0
    \text{ for every }\epsilon>0
    \text{ and every }\varphi
  \right\}.
  \label{eq:local-scaling-degree-definition}
\end{equation}
Finite scaling degree means \(\operatorname{sd}_0(T)<\infty\).  With this
normalization a smooth function has scaling degree \(0\), while
\(\partial^\alpha\delta_0\) in \(\mathbb R^n\) has scaling degree
\(n+|\alpha|\).
\end{definition}

\begin{proposition}[Microlocal status of the Dyson light-cone lift]
\label{prop:dyson-light-cone-lift-microlocal-criterion}
Let
\[
  X^\circ=(\mathbb M^4\times\mathbb R^2)\setminus\{(0,0)\},
  \qquad
  g(x,y)=x_D^2-|y|^2,
  \qquad
  \pi(x,y)=x ,
\]
where \(x_D^2=(x^0)^2-|\vec x|^2\).  Let
\[
  \mathcal N_{\rm lc}
  =
  \{(x;\lambda\,\dd_x x_D^2)\mid x\ne0,\ x_D^2=0,\ \lambda\in\mathbb R^\times\}
  \subset T^*\mathbb M^4\setminus0
\]
be the conormal bundle of the punctured four-dimensional light cone.  If
\(C\in\mathcal D'(\mathbb M^4)\) satisfies
\begin{equation}
  \operatorname{WF}(C)\cap\mathcal N_{\rm lc}=\varnothing ,
  \label{eq:dyson-lift-wavefront-transversality}
\end{equation}
then the product
\begin{equation}
  U^\circ=\pi^*C\cdot\delta(g)
  \label{eq:dyson-lift-punctured-product}
\end{equation}
is a well-defined distribution on \(X^\circ\).  If
\(\operatorname{supp}C\subset\overline V_{\rm caus}\), then
\[
  \operatorname{supp}U^\circ
  \subset
  \{(x,y)\in X^\circ\mid x\in\overline V_{\rm caus},\ x_D^2=|y|^2\}.
\]
Any two extensions of \(U^\circ\) to
\(\mathbb M^4\times\mathbb R^2\) differ by a distribution supported at
\((0,0)\).  Hence their six-dimensional Fourier transforms differ by a
polynomial in the six dual momenta, and their restrictions to \(p_1=p_2=0\)
differ by a polynomial in \(q\).

Consequently, if
\(\operatorname{WF}(C)\) contains a nonzero covector conormal to the
punctured light cone, Dyson's displayed product
\[
  C(x)\delta(x_D^2-|y|^2)
\]
is not canonically defined by the standard product of distributions.  In that
case a Dyson lift, if it exists, contains extra renormalized extension data;
that data is invisible away from the light-cone vertex only after one proves
the corresponding finite-contact ambiguity and the six-dimensional wave
equation.
\end{proposition}

\begin{proof}
The projection \(\pi:X^\circ\to\mathbb M^4\) is a submersion, so the pullback
\(\pi^*C\) is always defined and
\begin{equation}
  \operatorname{WF}(\pi^*C)
  \subset
  \{(x,y;\xi,0)\mid (x;\xi)\in\operatorname{WF}(C)\}.
  \label{eq:dyson-pullback-wavefront}
\end{equation}
On \(X^\circ\), the hypersurface \(g=0\) is smooth: the only zero of
\(\dd g\) is \((0,0)\), which has been removed.  Therefore
\begin{equation}
  \operatorname{WF}(\delta(g))
  =
  \{(x,y;\lambda\,\dd_x g,\lambda\,\dd_y g)
    \mid g(x,y)=0,\ \lambda\in\mathbb R^\times\}.
  \label{eq:dyson-delta-wavefront}
\end{equation}
Hormander's product criterion says that the product of two distributions is
defined if their wavefront sets contain no pair of opposite covectors over
the same base point.  A possible obstruction to multiplying
\(\pi^*C\) by \(\delta(g)\) would therefore require
\[
  (\xi,0)+(\lambda\,\dd_x g,\lambda\,\dd_y g)=0 .
\]
If \(y\ne0\), then \(\dd_y g=-2y\cdot\dd y\ne0\), and the second component
cannot vanish for \(\lambda\ne0\).  Hence the only obstruction can occur at
\(y=0\).  On \(g=0\), this forces \(x_D^2=0\), and the remaining obstruction
is precisely
\[
  (x;\xi)=
  (x;-\lambda\,\dd_x x_D^2)
  \in\mathcal N_{\rm lc}.
\]
The transversality condition
\eqref{eq:dyson-lift-wavefront-transversality} excludes this case and proves
that \eqref{eq:dyson-lift-punctured-product} is a well-defined distribution
on \(X^\circ\).

The support statement follows from the elementary support rule for a defined
product of distributions:
\[
  \operatorname{supp}(\pi^*C\cdot\delta(g))
  \subset
  \pi^{-1}(\operatorname{supp}C)\cap\{g=0\}.
\]
If \(U_1\) and \(U_2\) are two extensions to the full six-dimensional space,
then \(U_1-U_2\) vanishes on \(X^\circ\), so its support is contained in the
single point \((0,0)\).  A distribution supported at one point is a finite
linear combination of derivatives of the delta distribution at that point:
\[
  U_1-U_2
  =
  \sum_{|\alpha|+|\beta|\le N}
  c_{\alpha\beta}\,
  \partial_x^\alpha\partial_y^\beta
  \delta(x)\delta(y).
\]
Fourier transformation sends this finite sum to a polynomial in
\((q,p_1,p_2)\), and restriction to \(p_1=p_2=0\) gives a polynomial in
\(q\).  This is exactly the local-contact ambiguity recorded in the Dyson
representation.
\end{proof}

\begin{theorem}[Punctured Dyson lift and finite scaling degree]
\label{thm:punctured-dyson-lift-finite-scaling-degree}
Let \(C\in\mathcal S'(\mathbb M^4)\) be supported in
\(\overline V_{\rm caus}\) and suppose that the wavefront transversality
condition
\[
  \operatorname{WF}(C)\cap\mathcal N_{\rm lc}=\varnothing
\]
holds on the punctured four-dimensional light cone.  Let
\begin{equation}
  U^\circ=\kappa\,\pi^*C\cdot\delta(x_D^2-|y|^2)
  \qquad
  \text{on }
  X^\circ=(\mathbb M^4\times\mathbb R^2)\setminus\{0\},
  \label{eq:punctured-dyson-lift-Ucirc}
\end{equation}
where \(\kappa\ne0\) is a fixed normalization.  Assume that
\(U^\circ\) has finite scaling degree at the origin in
\(\mathbb R^6\), and that it admits a tempered \(O(2)\)-invariant extension
\[
  U\in\mathcal S'(\mathbb R^6)
\]
with the following three further properties:
\begin{equation}
  \operatorname{supp}U
  \subset
  \{(x,y)\mid x\in\overline V_{\rm caus},\ x_D^2=|y|^2\}.
  \label{eq:dyson-extension-support}
\end{equation}
\begin{equation}
  (x_D^2-|y|^2)\,U=0,
  \label{eq:dyson-extension-cone-equation}
\end{equation}
and the \(y\)-pushforward of \(U\), with the chosen normalization \(\kappa\),
equals \(C\) modulo a distribution supported at \(x=0\).
Then the Fourier transform
\[
  F(q,p_1,p_2)=\widehat U(q,p_1,p_2)
\]
is rotationally invariant in \((p_1,p_2)\) and solves
\begin{equation}
  \Box_{1,5}F=0
  \label{eq:dyson-extension-wave-equation}
\end{equation}
in the distributional sense.  The restriction of \(F\) to \(p_1=p_2=0\), if
defined by the chosen extension, equals \(\widehat C(q)\) up to a polynomial
in \(q\).  Hence \(C\) is Dyson-regular modulo contact terms once the
restriction and Cauchy-data requirements in
Definition~\ref{def:dyson-regular-causal-commutator} are satisfied by \(F\).

Moreover, if \(U_1\) and \(U_2\) are two such finite-scaling-degree
extensions, then
\begin{equation}
  U_1-U_2
  =
  \sum_{|\alpha|+|\beta|\le N}
  c_{\alpha\beta}\,
  \partial_x^\alpha\partial_y^\beta\delta(x)\delta(y)
  \label{eq:dyson-extension-local-ambiguity}
\end{equation}
for some \(N\).  Thus the induced ambiguity in \(F(q,0,0)\) is a finite
polynomial in \(q\), and no cut, threshold, or hyperboloid support is changed
by choosing a different extension.
\end{theorem}

\begin{proof}
The product \(U^\circ\) on \(X^\circ\) is well-defined by
Proposition~\ref{prop:dyson-light-cone-lift-microlocal-criterion}.  We first
recall the extension mechanism because it is the local analytic part of the
argument.  Let \(n=6\) and let \(T^\circ\in\mathcal D'(\mathbb R^n\setminus
\{0\})\) have finite scaling degree \(s\) at the origin.  Put
\[
  m=\lfloor s\rfloor-n
\]
when \(s\ge n\), and \(m=-1\) otherwise.  Choose
\(\chi\in C_c^\infty(\mathbb R^n)\) equal to \(1\) near the origin.  For a
test function \(\varphi\), define
\begin{equation}
  R_m\varphi(z)
  =
  \varphi(z)
  -
  \chi(z)
  \sum_{|\alpha|\le m}
  \frac{z^\alpha}{\alpha!}\partial^\alpha\varphi(0)
  \label{eq:dyson-scaling-degree-taylor-remainder}
\end{equation}
with the sum omitted when \(m=-1\).  The function \(R_m\varphi\) vanishes at
the origin to order \(m+1\).  Let \(\chi_\epsilon(z)=\chi(z/\epsilon)\).
The finite-scaling-degree bound makes the limit
\[
  T_0(\varphi)
  =
  \lim_{\epsilon\downarrow0}
  T^\circ\bigl((1-\chi_\epsilon)R_m\varphi\bigr)
\]
finite and continuous in the test-function topology: the Taylor remainder
removes precisely the jets whose scaling would make the punctured
distribution divergent at the origin.  The possible values of an extension on
the finite jet removed in
\eqref{eq:dyson-scaling-degree-taylor-remainder} are exactly the constants
multiplying \(\partial^\alpha\delta_0\) for \(|\alpha|\le m\).  This gives an
extension of \(T^\circ\), preserves the scaling degree, and shows that two
extensions with the same punctured distribution differ by a finite sum of
derivatives of \(\delta_0\).  If \(T^\circ\) is supported in a closed cone
with vertex at the origin, the same construction applied after restricting
test functions supported away from that cone gives an extension with support
in the closed cone; the only new support allowed by extension is the vertex
itself.

Applying this statement to \(T^\circ=U^\circ\) gives the asserted local
ambiguity \eqref{eq:dyson-extension-local-ambiguity}.  The additional
condition \eqref{eq:dyson-extension-cone-equation} is essential: support on
the cone alone would still allow derivatives of the delta distribution in
the normal direction at the vertex, and multiplication by the quadratic
equation need not annihilate such terms.  Fourier transformation converts
multiplication by \(x_D^2-|y|^2\) into the six-dimensional wave operator in
momentum space, up to the sign fixed by the Fourier convention.  Therefore
\(\Box_{1,5}F=0\).  Rotational invariance in the two auxiliary momenta is one
of the stated extension requirements.  Equivalently, if a non-invariant
extension satisfying the other conditions has been constructed, averaging it
over \(O(2)\) preserves the punctured value, support,
\eqref{eq:dyson-extension-cone-equation}, and the pushforward normalization.

It remains to identify the four-dimensional restriction.  On
\(X^\circ\), the \(y\)-pushforward of the punctured expression is the intended
four-dimensional commutator with the normalization \(\kappa\).  The theorem
therefore requires the extension to preserve this pushforward modulo a
distribution supported at \(x=0\).  By
Proposition~\ref{prop:auxiliary-pushforward-momentum-trace}, Fourier
transformation of this \(y\)-pushforward is precisely the \(p=0\) trace of
\(\widehat U(q,p)\), when either side exists in the stated distributional
sense.  Hence the distribution obtained by restricting \(\widehat U(q,p)\)
to \(p=0\) differs from \(\widehat C(q)\) by the Fourier transform of a
distribution supported at \(x=0\), namely by a polynomial in \(q\).  The same
conclusion follows directly from
\eqref{eq:dyson-extension-local-ambiguity}: setting \(p=0\) turns the
Fourier transform of every
\(\partial_x^\alpha\partial_y^\beta\delta(x)\delta(y)\) into a polynomial in
\(q\), with the \(p\)-powers evaluated at zero.

The theorem has not proved the physical source-current kernels to be
Dyson-regular.  It proves the exact local construction that remains once the
punctured wavefront condition, finite scaling degree, cone equation,
pushforward normalization, temperedness, and Cauchy-data trace hypotheses
have been verified for those kernels.  Under those hypotheses the difference
between the constructed lift and a strict Dyson lift is only a contact term,
so
Definition~\ref{def:dyson-regular-modulo-contact} applies.
\end{proof}

\begin{proposition}[Source-current differentials preserve the punctured input]
\label{prop:source-current-differentials-preserve-punctured-input}
Let \(\Phi_b,\Phi_c\) be local scalar fields and let
\(\Psi_i,\Psi_f\) be normalizable finite-energy vectors.  Define the
relative-coordinate commutator matrix element
\begin{equation}
  C^\Phi_{bc}(x)
  =
  \langle\Psi_f,[\Phi_b(x/2),\Phi_c(-x/2)]\Psi_i\rangle
  \in\mathcal S'(\mathbb M^4).
  \label{eq:field-commutator-relative-coordinate}
\end{equation}
Let \(\mathcal K_b(\partial_{x_b})\) and
\(\mathcal K_c(\partial_{x_c})\) be the mostly-plus Klein--Gordon LSZ
operators used to define source currents
\[
  J_b=\mathcal K_b\Phi_b,\qquad J_c=\mathcal K_c\Phi_c.
\]
Set
\begin{equation}
  P_{bc}(\partial_x)
  =
  \mathcal K_b(2\partial_x)\mathcal K_c(-2\partial_x),
  \qquad
  r=\operatorname{ord}P_{bc}.
  \label{eq:source-current-relative-differential-operator}
\end{equation}
Then
\begin{equation}
  C^J_{bc}(x)
  =
  \langle\Psi_f,[J_b(x/2),J_c(-x/2)]\Psi_i\rangle
  =
  P_{bc}(\partial_x)C^\Phi_{bc}(x)
  \label{eq:source-current-commutator-differential-reduction}
\end{equation}
as tempered distributions.  Consequently:
\begin{enumerate}
  \item If \(\operatorname{supp}C^\Phi_{bc}\subset\overline V_{\rm caus}\),
  then
  \[
    \operatorname{supp}C^J_{bc}\subset\overline V_{\rm caus}.
  \]
  \item If
  \[
    \operatorname{WF}(C^\Phi_{bc})\cap\mathcal N_{\rm lc}=\varnothing ,
  \]
  then
  \[
    \operatorname{WF}(C^J_{bc})\cap\mathcal N_{\rm lc}=\varnothing .
  \]
  \item If \(C^\Phi_{bc}\) has finite scaling degree \(s\) at \(x=0\), then
  \(C^J_{bc}\) has finite scaling degree at most \(s+r\).
  \item Under the preceding wavefront and scaling assumptions, the punctured
  Dyson product
  \begin{equation}
    U^{J,\circ}_{bc}
    =
    \kappa\,\pi^*C^J_{bc}\cdot\delta(x_D^2-|y|^2)
    \qquad\text{on }X^\circ
    \label{eq:source-current-punctured-dyson-product}
  \end{equation}
  is well-defined and has finite scaling degree at the six-dimensional
  origin, bounded by \(s+r+2\).
\end{enumerate}
Thus replacing interpolating fields by LSZ source currents supplies no new
punctured light-cone obstruction beyond finite-order differentiation.  The
remaining, genuinely nonlocal input for
Theorem~\ref{thm:punctured-dyson-lift-finite-scaling-degree} is the existence
of a tempered \(O(2)\)-invariant extension satisfying the cone equation,
pushforward normalization, and Cauchy-data trace requirements.
\end{proposition}

\begin{proof}
The coordinate \(x_b=x/2\) gives
\[
  \partial_{x_b^\mu}=2\partial_{x^\mu},
\]
whereas \(x_c=-x/2\) gives
\[
  \partial_{x_c^\mu}=-2\partial_{x^\mu}.
\]
Since the LSZ source currents are obtained by constant-coefficient
differential operators acting on the corresponding field variable, the
commutator matrix element is obtained from
\eqref{eq:field-commutator-relative-coordinate} by the operator
\eqref{eq:source-current-relative-differential-operator}.  This proves
\eqref{eq:source-current-commutator-differential-reduction}.

Differentiation cannot enlarge support, so causal support is preserved.
For wavefront sets, every constant-coefficient differential operator \(P\)
satisfies
\begin{equation}
  \operatorname{WF}(PT)\subset\operatorname{WF}(T).
  \label{eq:differential-wavefront-inclusion}
\end{equation}
Indeed, after multiplying by a cutoff \(\chi\) equal to one near a point, the
Fourier transform of \(\chi PT\) is a polynomial in the Fourier variable
times the Fourier transform of \(\chi T\), plus lower-order terms with
cutoffs supported in the same neighbourhood; polynomial multiplication cannot
destroy rapid decay in a conic neighbourhood where \(\chi T\) already has
rapid decay.  This proves the wavefront assertion.

The scaling-degree estimate is equally direct.  For a multi-index \(\alpha\),
\begin{equation}
  \langle\partial^\alpha T,\varphi_\lambda\rangle
  =
  (-1)^{|\alpha|}
  \lambda^{-|\alpha|}
  \langle T,(\partial^\alpha\varphi)_\lambda\rangle .
  \label{eq:scaling-degree-derivative-shift}
\end{equation}
Thus \(\operatorname{sd}_0(\partial^\alpha T)\le
\operatorname{sd}_0(T)+|\alpha|\).  Applying this to every derivative in
\(P_{bc}\) gives \(\operatorname{sd}_0(C^J_{bc})\le s+r\).

The punctured product
\eqref{eq:source-current-punctured-dyson-product} is defined by
Proposition~\ref{prop:dyson-light-cone-lift-microlocal-criterion}, because
the source-current commutator has the same conormal avoidance.  It remains
only to record finite scaling degree in six dimensions.  The pullback
\(\pi^*C^J_{bc}\) has the same local scaling degree as \(C^J_{bc}\): when a
six-dimensional test function is scaled, integration over the two auxiliary
\(y\)-variables produces a scaled four-dimensional test function with the
same power of \(\lambda\) relevant to the \(x\)-distribution.  The
distribution \(\delta(x_D^2-|y|^2)\) is homogeneous of degree \(-2\), hence
has scaling degree \(2\) in \(\mathbb R^6\).  The product is a continuous
microlocal product on the conic wavefront neighbourhoods fixed in the
previous proposition, and the defining seminorm estimates give the standard
subadditivity bound
\begin{equation}
  \operatorname{sd}_0
  \bigl(\pi^*C^J_{bc}\cdot\delta(x_D^2-|y|^2)\bigr)
  \le
  \operatorname{sd}_0(C^J_{bc})+2
  \le s+r+2.
  \label{eq:source-current-lift-scaling-bound}
\end{equation}
This proves the final assertion.
\end{proof}

\begin{theorem}[Dyson representation of a causal commutator]
\label{thm:dyson-causal-commutator-representation}
Let \(C\in\mathcal S'(\mathbb M^4)\) be Dyson-regular in the sense of
Definition~\ref{def:dyson-regular-causal-commutator}.  Assume
\(\operatorname{supp}C\subset\overline V_{\rm caus}\).  Then there is a
distribution
\[
  \Psi\in\mathcal D'(\mathbb R^4_u\times[0,\infty)_\mu)
\]
such that, in \(\mathcal D'(\mathbb R^4_q)\),
\begin{equation}
  \widehat C(q)
  =
  \left\langle
    \Psi(u,\mu),
    \varepsilon(q^0-u^0)\,
    \delta\!\bigl((q-u)^2+\mu^2\bigr)
  \right\rangle_{u,\mu}.
  \label{eq:dyson-causal-commutator-representation}
\end{equation}
Here \(\varepsilon(t)=1\) for \(t>0\) and \(\varepsilon(t)=-1\) for
\(t<0\).  If, in addition, \(\widehat C\) vanishes in the slab \(R\) of
\eqref{eq:dyson-coincidence-slab} and the Dyson data are support-pruned
relative to \(R\) in the sense of
Definition~\ref{def:support-pruned-dyson-data}, then \(\Psi\) can be chosen
with
\[
  \operatorname{supp}\Psi\subset S_R .
\]
If a later source-current or retarded-product reduction adds local terms
supported at \(x=0\), their Fourier transforms are finite polynomials in
\(q\).  Such local-contact polynomials are recorded separately; they do not
change the hyperboloid support or the holomorphic domain away from infinity.
\end{theorem}

\begin{proof}
The proof is a hyperbolic Cauchy problem in disguise.  We give it because the
support statement is the part that carries the QFT locality information into
the analytic domain.  The Dyson-regularity hypothesis supplies the
six-dimensional wave-equation solution \(F\); without that hypothesis, the
formal product \(C(x)\delta(x_D^2-|y|^2)\) at the light cone would not be a
legitimate operation for an arbitrary tempered distribution.

Temporarily write the scalar product in Dyson's mostly-minus convention
\[
  x_D^2=(x^0)^2-|\vec x|^2=-x^2 .
\]
The causal support hypothesis becomes \(C(x)=0\) for \(x_D^2<0\).  Add two
auxiliary spacelike coordinates \(y=(y_1,y_2)\) and set
\[
  Z=(x,y)\in\mathbb R^{1,5},
  \qquad
  Z_D^2=x_D^2-y_1^2-y_2^2 .
\]
Let \(R_6=(q,p_1,p_2)\) be the six-dimensional momentum dual to \(Z\).  By
Dyson regularity there is a rotationally invariant distribution \(F(R_6)\)
such that
\begin{equation}
  \Box_{1,5}F=0 .
  \label{eq:dyson-six-dimensional-wave-equation}
\end{equation}
and the original Fourier transform is recovered as the boundary value
\[
  \widehat C(q)=F(q,0,0)
\]
after absorbing the fixed normalization constant in the definition of the
lift.  In the regular-function case this \(F\) is the Fourier transform of
\(4\pi C(x)\delta(Z_D^2)\), which explains the geometry of the construction.

For solutions of \eqref{eq:dyson-six-dimensional-wave-equation} with the
Dyson-regular Cauchy traces,
Proposition~\ref{prop:six-dimensional-distributional-cauchy-formula} gives,
on every affine spacelike hyperplane
\(\Sigma\subset\mathbb R^{1,5}\),
\begin{equation}
  F(R_0)
  =
  \int_\Sigma
  \left(
    F(R)\,\partial_{n_R}D_6(R-R_0)
    -
    D_6(R-R_0)\,\partial_{n_R}F(R)
  \right)\dd\Sigma(R),
  \label{eq:dyson-cauchy-wave-formula}
\end{equation}
where \(D_6\) is the odd fundamental solution of the six-dimensional wave
operator and \(n_R\) is the future unit normal to \(\Sigma\).  This is a
theorem about the linear wave equation and its distributional Cauchy data; no
QFT input enters it after Dyson regularity has supplied the traces.

Take \(R_0=(q,0,0)\).  Because \(D_6\) is supported on the
six-dimensional light cone, and because \(F\) is rotation-invariant in the
last two momenta, the angular integral over the \((p_1,p_2)\)-plane rewrites
\eqref{eq:dyson-cauchy-wave-formula} as a superposition of four-dimensional
mass hyperboloids:
\begin{equation}
  F(q,0,0)
  =
  \left\langle
    \Psi_\Sigma(u,\mu),
    \varepsilon(q_D^0-u_D^0)
    \delta\!\bigl((q_D-u_D)^2-\mu^2\bigr)
  \right\rangle .
  \label{eq:dyson-mostly-minus-hyperboloid-representation}
\end{equation}
Here \(u\) is the four-momentum part of the point \(R\in\Sigma\), while
\(\mu=(p_1^2+p_2^2)^{1/2}\); the distribution \(\Psi_\Sigma\) is the push
forward of the Cauchy data
\((F|_\Sigma,\partial_nF|_\Sigma)\) with the differential operator that comes
from \(\partial_nD_6\).  Returning to the mostly-plus convention changes
\((q_D-u_D)^2-\mu^2=0\) into
\((q-u)^2+\mu^2=0\), giving the kernel in
\eqref{eq:dyson-causal-commutator-representation}.

The support-restricted conclusion is now exactly the content of
Definition~\ref{def:support-pruned-dyson-data}.  If the commutator vanishes
in the slab \(R\) and its Dyson data are support-pruned relative to \(R\),
choose \(\Psi_R\) satisfying
\eqref{eq:support-pruned-dyson-data}.  Since
\(\mathcal H\Psi_R=\mathcal H\Psi_\Sigma=\widehat C\), the same
hyperboloid representation holds with
\(\operatorname{supp}\Psi_R\subset S_R\).

The geometric reason for the support-pruning condition is the following.
If \((u,\mu)\notin S_R\), at least one sheet of the hyperboloid
\((q-u)^2=-\mu^2\) enters the coincidence slab.  Since the slab boundaries
are spacelike or null graphs, a small double cone in momentum space adapted
to that sheet lies inside \(R\).  The classical Jost--Lehmann--Dyson proof
uses the six-dimensional wave equation and domain-of-dependence uniqueness
to remove the corresponding Cauchy-data component.  In this monograph that
removal is recorded as the support-pruning theorem still to be proved for
the LSZ source-current kernels; it is not identified with the simpler
statement that the trace \(F(q,0,0)\) vanishes in \(R\).

The statement about finite contact polynomials follows from differentiating
the commutator or from derivatives of step functions in retarded products:
Fourier transformation changes local distributions supported at \(x=0\) into
polynomials in \(q\).  Such terms have no cut or hyperboloid support and are
separated as subtraction/contact terms in dispersion formulae.
\end{proof}

\begin{proposition}[Coincidence regions from the spectral condition]
\label{prop:source-current-commutator-coincidence-region}
Let \(J_b,J_c\) be local source-current fields whose matrix elements with the
states below define tempered distributions.  Let \(\Psi_i,\Psi_f\) be
normalizable vectors whose joint energy-momentum spectral supports are compact
sets \(B_i,B_f\subset\operatorname{Spec}(P)\), and let \(E(\cdot)\) be the
joint spectral measure of \(P^\mu\).  Define the intermediate spectral
matrix distributions
\begin{align}
  \dd\nu^+_{bc}(r_f,p,r_i)
  &=
  \langle E(\dd r_f)\Psi_f,
    J_b(0)E(\dd p)J_c(0)E(\dd r_i)\Psi_i\rangle,
  \nonumber\\
  \dd\nu^-_{bc}(r_f,p,r_i)
  &=
  \langle E(\dd r_f)\Psi_f,
    J_c(0)E(\dd p)J_b(0)E(\dd r_i)\Psi_i\rangle .
  \label{eq:source-current-intermediate-spectral-distributions}
\end{align}
Let \(\mathfrak S^+_{bc}\) and \(\mathfrak S^-_{bc}\) be their supports in
\[
  B_f\times\operatorname{Spec}(P)\times B_i .
\]
With the
Fourier convention
\[
  \widehat f(q)=\int\ee^{\ii q\cdot x}f(x)\,\dd^4x,
\]
the two ordered matrix elements
\[
  C^+_{bc}(x)=
  \langle\Psi_f,J_b(x/2)J_c(-x/2)\Psi_i\rangle,
  \qquad
  C^-_{bc}(x)=
  \langle\Psi_f,J_c(-x/2)J_b(x/2)\Psi_i\rangle
\]
obey
\begin{align}
  \operatorname{supp}\widehat C^+_{bc}
  \subset
  \left\{
    p-\frac12(r_f+r_i)
    \ \middle|\ (r_f,p,r_i)\in\mathfrak S^+_{bc}
  \right\},
  \nonumber\\
  \operatorname{supp}\widehat C^-_{bc}
  \subset
  \left\{
    \frac12(r_f+r_i)-p
    \ \middle|\ (r_f,p,r_i)\in\mathfrak S^-_{bc}
  \right\}.
  \label{eq:source-current-ordered-supports-wavepacket}
\end{align}
Hence the commutator distribution
\[
  C_{bc}=C^+_{bc}-C^-_{bc}
\]
has Fourier transform supported in the union of the two displayed sets.
Every open set \(R\) disjoint from this union is a coincidence region:
\begin{equation}
  \widehat C_{bc}(q)=0\qquad(q\in R).
  \label{eq:source-current-coincidence-vanishing}
\end{equation}
The coarser but often sufficient replacement
\[
  \mathfrak S^\pm_{bc}\subset
  B_f\times\overline V_+\times B_i
\]
uses
\[
  K_{fi}=\frac12(B_f+B_i)
  =
  \left\{\frac12(r_f+r_i)\ \middle|\ r_f\in B_f,\ r_i\in B_i\right\}
\]
and gives the model coincidence region between the two shifted spectral cones
\[
  \overline V_+-K_{fi},
  \qquad
  K_{fi}-\overline V_+ .
\]
For generalized sharp external momenta \(Q_i,Q_f\), one recovers the familiar
shorthand by setting \(K=(Q_f+Q_i)/2\) and
\[
  \Sigma_{bc}^{fi}=\operatorname{supp}
  \langle J_b(0)^*\Psi_f,E(\dd p)J_c(0)\Psi_i\rangle,
  \qquad
  \Sigma_{cb}^{fi}=\operatorname{supp}
  \langle J_c(0)^*\Psi_f,E(\dd p)J_b(0)\Psi_i\rangle ,
\]
so that the two ordered supports are contained in
\[
  \Sigma_{bc}^{fi}-K,\qquad K-\Sigma_{cb}^{fi}.
\]
\end{proposition}

\begin{proof}
Translation covariance gives
\[
  J_b(x/2)=\ee^{\ii P\cdot x/2}J_b(0)\ee^{-\ii P\cdot x/2},
  \qquad
  J_c(-x/2)=\ee^{-\ii P\cdot x/2}J_c(0)\ee^{\ii P\cdot x/2}.
\]
Insert the spectral resolution of \(P\) on the external vectors and between
the two currents.  The first ordered matrix element becomes
\begin{equation}
\begin{aligned}
  C^+_{bc}(x)
  &=
  \int
  \ee^{\ii(r_f+r_i)\cdot x/2}
  \ee^{-\ii p\cdot x}\,
  \dd\nu^+_{bc}(r_f,p,r_i)\\
  &=
  \int
  \ee^{\ii((r_f+r_i)/2-p)\cdot x}\,
  \dd\nu^+_{bc}(r_f,p,r_i),
\end{aligned}
  \label{eq:source-current-first-ordered-spectral-form}
\end{equation}
where the equality is first read after smearing all fields and then extended
to the matrix-element distribution by Wightman temperedness.  Pairing against
a test function in \(x\) and using the Fourier convention in the statement
shows that \(\widehat C^+_{bc}\) is supported where
\[
  q+\frac12(r_f+r_i)-p=0.
\]
Thus \(q=p-\frac12(r_f+r_i)\) with
\((r_f,p,r_i)\in\mathfrak S^+_{bc}\).  This is the first inclusion in
\eqref{eq:source-current-ordered-supports-wavepacket}.

The second ordered product is
\begin{equation}
\begin{aligned}
  C^-_{bc}(x)
  &=
  \int
  \ee^{-\ii(r_f+r_i)\cdot x/2}
  \ee^{\ii p\cdot x}\,
  \dd\nu^-_{bc}(r_f,p,r_i)\\
  &=
  \int
  \ee^{\ii(p-(r_f+r_i)/2)\cdot x}\,
  \dd\nu^-_{bc}(r_f,p,r_i).
\end{aligned}
  \label{eq:source-current-second-ordered-spectral-form}
\end{equation}
Therefore its Fourier transform is supported where
\[
  q+p-\frac12(r_f+r_i)=0,
\]
or \(q=\frac12(r_f+r_i)-p\), proving the second inclusion.  Taking the
difference of the two ordered products proves the support inclusion for the
commutator.  A distribution whose Fourier support is contained in a closed set
vanishes on every open set disjoint from that support, which proves
\eqref{eq:source-current-coincidence-vanishing}.
\end{proof}

\begin{remark}[JLD input for the LSZ retarded commutator]
\label{rem:jld-input-for-lsz-retarded-commutator}
Let
\[
  J_b=\mathcal K_b\Phi_b,\qquad J_c=\mathcal K_c\Phi_c
\]
denote the source currents obtained by applying the mostly-plus
Klein--Gordon operators \(\mathcal K_b,\mathcal K_c\) used in the LSZ
reduction formula.  For finite-energy states \(|Q_i\rangle,|Q_f\rangle\),
assume that the distribution
\[
  C_{bc}(x)
  =
  \langle Q_f,[J_b(x/2),J_c(-x/2)]Q_i\rangle
\]
is Dyson-regular modulo contact terms in the sense of
Definition~\ref{def:dyson-regular-modulo-contact}.  Then the nonlocal part of
\(C_{bc}\) has a Dyson representation of the form
\eqref{eq:dyson-causal-commutator-representation}, plus a finite polynomial in
\(q\) from contact terms.  In a coincidence slab \(R\) where the
intermediate-state spectral condition makes the nonlocal part of
\(\widehat C_{bc}\) vanish, this representation is support-restricted to
\(S_R\) if and only if the corresponding Dyson data are support-pruned
relative to \(R\) in the sense of
Definition~\ref{def:support-pruned-dyson-data}.  The associated retarded and
advanced Fourier
transforms
\[
  C^R_{bc}(q)=\int\ee^{\ii q\cdot x}\theta(x^0)C_{bc}(x)\,\dd^4x,
  \qquad
  C^A_{bc}(q)=-\int\ee^{\ii q\cdot x}\theta(-x^0)C_{bc}(x)\,\dd^4x
\]
are boundary values of the same analytic continuation across that
coincidence region, up to the finite contact polynomials generated by the
source-current reduction.
The justification is exactly the chain of hypotheses now separated above.
Locality gives causal support; Proposition~\ref{prop:source-current-commutator-coincidence-region}
gives the spectral coincidence regions; Dyson regularity is the extra
six-dimensional wave-equation input needed to use
Theorem~\ref{thm:dyson-causal-commutator-representation}; support restriction
inside a slab is a property of the Cauchy-data class modulo
\(\ker\mathcal H\), together with the separate datum that
\(\widehat C_{bc}=0\) on the slab.  Local source-current terms supported at
\(x=0\) transform to finite
polynomials in \(q\).  The edge-of-the-wedge continuation of retarded and
advanced boundary values is therefore a conditional analytic consequence of
these inputs, with theorem status supplied only after the hypotheses are
verified.
\end{remark}

Remark~\ref{rem:primitive-domain-output-for-scattering} is the rigorous
crossing-domain input supplied directly by the Wightman locality and spectrum
axioms, while
Proposition~\ref{prop:source-current-commutator-coincidence-region} supplies
the spectral coincidence regions for the LSZ source-current commutator.
Theorem~\ref{thm:dyson-causal-commutator-representation} then supplies the
causal-commutator representation used to enlarge domains from those
coincidence regions, provided the commutator is Dyson-regular modulo contact
terms; the \(S_R\)-support-restricted version additionally requires
support-pruned Dyson data relative to the coincidence slab.  The stronger
fixed-\(t\) cut-plane theorem used in dispersion relations still requires the
Dyson-regularity and support-pruning verification for the relevant LSZ
source-current matrix coefficients.
Proposition~\ref{prop:dyson-light-cone-lift-microlocal-criterion} shows the
precise microlocal obstruction to treating Dyson's light-cone lift as an
ordinary product of distributions, and identifies the finite local ambiguity
of any renormalized lift.
Theorem~\ref{thm:punctured-dyson-lift-finite-scaling-degree} then proves the
local finite-scaling-degree extension mechanism, including the cone equation,
pushforward normalization, and the fact that all remaining extension
ambiguity is polynomial after Fourier transformation.
Proposition~\ref{prop:source-current-differentials-preserve-punctured-input}
checks that passing from interpolating fields to LSZ source currents preserves
the punctured wavefront and finite-scaling-degree inputs by finite-order
differentiation, with the correct \(x/2\) and \(-x/2\) factors.  What remains
is not the differential source-current step, but the actual QFT theorem
constructing the required extension, Cauchy data, and support-pruning for the
source-current kernel.  After that regularity problem one still needs the
Bros--Epstein--Glaser analytic-completion step and the LSZ
pole-separation transfer hypotheses.  The analytic LSZ transfer itself is
proved in
Theorem~\ref{thm:lsz-transfer-fixed-t-domain}: once an off-shell fixed-\(t\)
domain has normal-crossing external one-particle poles with uniform boundary
growth, restricting the LSZ-amputated germ to the external mass shells
preserves holomorphy and tempered discontinuities.  What remains here is to
prove that the primitive four-point germ and the causal-commutator Dyson
representation extend far enough to give such an off-shell first-sheet
\(s\)-plane domain at fixed \(t\), with only the stated right- and left-hand
cuts and with controlled boundary growth.  That additional step is the
Lehmann--Martin/Jin--Martin domain theorem, not a restatement of primitive
tube analyticity.

The same symbol \(\mathcal M(s,t)\) will denote the holomorphic germ
\(F(s,t,4m^2-s-t)\) on whichever connected domain is under discussion, and
its physical value will always mean the specified tempered boundary value.
Analytic continuation is continuation of this germ through overlapping
domains in \(\mathcal O_{\mathrm{poly}}\).  If two such germs have the same
tempered boundary value on a nonempty real open edge, the edge-of-the-wedge
theorem gives a common continuation to the envelope generated by the adjacent
tubes; inside a connected holomorphic domain the ordinary identity theorem
then gives uniqueness.  No perturbative Borel summation procedure is part of
this definition.  The polynomial growth in
\(\mathcal O_{\mathrm{poly}}\) is the local growth condition needed for
distributional boundary values; polynomial boundedness in high-energy
directions is a stronger global estimate and is introduced only in the
dispersion-relation chapter.

Hermiticity of the theory and the Feynman boundary prescription imply real
analyticity on conjugate domains:
\[
  \mathcal M(s^*,t^*)=\mathcal M(s,t)^*
\]
where both sides are evaluated on corresponding branches.  On a real segment
not lying on a cut, this says that the first-sheet value is real up to pole
singularities.  On a cut, it relates the lower-edge and upper-edge boundary
values.  In perturbation theory the same property is visible graph by graph
when the couplings are real.

Crossing states that the \(s\)-, \(t\)-, and \(u\)-channel physical amplitudes
are boundary values of the same analytically continued four-point function in
different real regions, with the external representation conjugated when an
outgoing particle is moved to the incoming side or conversely.  On the real
\((s,t)\)-plane these regions are separated by thresholds and cuts.  The
locality-and-spectrum argument gives a common analytic germ in the
Bros--Epstein--Glaser envelope generated by the relevant tube domains and
spacelike edge identifications.  A stronger statement, namely that full
physical \(s\)- and crossed-channel regions can be joined by a path on a
specified first-sheet domain, requires an analyticity domain large enough to
contain that path.  In four spacetime dimensions this larger Mandelstam-type
domain requires input beyond the edge-of-the-wedge argument alone.

The Mandelstam representation is a stronger conjectural statement about this
same four-point function.  For equal external masses, with
\(u=4m^2-s-t\), it would assert that, after all first-sheet pole terms and a
specified finite polynomial subtraction have been separated, the remaining
amplitude admits a simultaneous two-variable dispersion representation of the
form
\[
\begin{aligned}
  \mathcal M_{\rm reg}(s,t)
  &=
  \sum_{\chi=s,t,u}
  \int_{\chi_0}^{\infty}
  \frac{\rho_\chi(\sigma)}{\sigma-\chi}\,\dd\sigma
\\
  &\quad+
  \sum_{(\chi,\chi')=(s,t),(s,u),(t,u)}
  \int_{\chi_0}^{\infty}\!\dd\sigma
  \int_{\chi'_0}^{\infty}\!\dd\tau\,
  \frac{\rho_{\chi\chi'}(\sigma,\tau)}
  {(\sigma-\chi)(\tau-\chi')} ,
\end{aligned}
\]
with \(i0\) boundary prescriptions assigned when a physical edge of a cut is
approached.  Here \(\rho_\chi\) and \(\rho_{\chi\chi'}\) are single and double
discontinuity distributions, and the support restrictions on the double
spectral distributions are part of the conjectural content.  This
simultaneous-cut representation is stronger than the single-variable
Cauchy representation at fixed \(t\).  No derivation is presently available
for general interacting four-dimensional Wightman theories with a mass gap.
The arguments in this monograph use fixed-\(t\) or fixed-crossed-variable
dispersion relations together with crossing and angular analyticity; the
Mandelstam representation is not assumed in those deductions.

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.86]
  \draw[->,line width=0.75pt] (-3.7,0)--(3.85,0) node[right] {\(s\)};
  \draw[->,line width=0.75pt] (0,-3.0)--(0,3.15) node[above] {\(t\)};
  \coordinate (S) at (1.35,0);
  \coordinate (T) at (0,1.35);
  \node[below] at (S) {\(4m^2\)};
  \node[left] at (T) {\(4m^2\)};
  \draw[gray!55,line width=0.7pt] (-1.6,2.95)--(3.35,-2.0)
    node[below right] {\(u=0\)};
  \fill[blue!18] (S)--(3.35,0)--(3.35,-2.0)--cycle;
  \draw[blue!70!black,line width=1.35pt] (S)--(3.35,0);
  \draw[blue!70!black,line width=1.35pt] (S)--(3.35,-2.0);
  \node[blue!70!black,align=center,fill=white,inner sep=1.2pt]
    at (2.55,-0.75) {\(s\)-channel\\\(12\to34\)};
  \fill[green!18] (T)--(0,3.0)--(-1.65,3.0)--cycle;
  \draw[green!55!black,line width=1.35pt] (T)--(0,3.0);
  \draw[green!55!black,line width=1.35pt] (T)--(-1.65,3.0);
  \node[green!55!black,align=center,fill=white,inner sep=1.2pt]
    at (-0.72,2.25) {\(t\)-channel\\\(1\bar3\to\bar2 4\)};
  \fill[orange!18] (0,0)--(0,-3.0)--(-3.0,-3.0)--(-3.0,0)--cycle;
  \draw[orange!85!black,line width=1.35pt] (0,0)--(0,-3.0);
  \draw[orange!85!black,line width=1.35pt] (0,0)--(-3.0,0);
  \node[orange!85!black,align=center,fill=white,inner sep=1.2pt]
    at (-2.1,-1.15) {\(u\)-channel\\\(1\bar4\to3\bar2\)};
  \draw[purple!80!black,->,line width=1.05pt]
    (2.45,-0.18) .. controls (1.1,1.42) and (-0.55,1.76) .. (-1.05,2.02);
  \node[purple!80!black,align=center,fill=white,inner sep=1pt]
    at (0.45,1.18) {analytic\\continuation};
\end{tikzpicture}
\end{center}

\section{Unitarity on the Physical Cut}

Crossing and real analyticity constrain how different boundary values fit
together.  On the physical \(s\)-channel cut there is also the Hilbert-space
constraint from the unitary scattering operator.  Let \(S_\ell(s)\) denote the
angular-momentum-\(\ell\) partial-wave scattering eigenvalue in the
identical-scalar two-particle sector, with the normalization fixed earlier.
The elastic \(2\to2\) amplitude is
\[
  \mathcal M(s,t)
  =
  -16\pi\ii
  \sqrt{\frac{s}{s-4m^2}}
  \sum_{\ell=0}^{\infty}
  (2\ell+1)P_\ell(\cos\theta)\bigl(S_\ell(s)-1\bigr),
\]
where
\[
  \cos\theta=1+\frac{2t}{s-4m^2}.
\]
Unitarity of \(S\) implies, for \(s\ge4m^2\),
\[
  |S_\ell(s)|\le1 .
\]
If no inelastic channel is open, namely for
\[
  4m^2\le s\le M_{\mathrm{inel}}^2 ,
\]
the partial wave is a phase and
\[
  |S_\ell(s)|=1 .
\]
Here \(M_{\mathrm{inel}}\) is the invariant mass threshold of the lightest
state outside the two-particle elastic channel.
These inequalities are boundary conditions on the upper edge of the physical
cut; they are not by themselves an analytic continuation principle.  They will
combine with angular analyticity in the following chapter.

\section{Perturbative Discontinuities and Cut Rules}

The Hilbert-space unitarity statement determines the physical-cut
discontinuity of exact on-shell amplitudes whenever the relevant upper and
lower boundary values exist as distributions and the intermediate-state
resolution is complete in the scattering sector under consideration.  Its
perturbative expansion gives the graph-level cut rules.  Write
\[
  S=\mathbf 1+\ii T,
  \qquad
  \operatorname{Disc}_s\mathcal M
  =
  \mathcal M(s+\ii0,t)-\mathcal M(s-\ii0,t).
\]
From \(S^\dagger S=\mathbf 1\) one obtains
\[
  -\ii(T-T^\dagger)=T^\dagger T .
\]
Taking connected matrix elements and removing the common momentum-conserving
delta function gives, on the \(s\)-channel physical cut,
\begin{equation}
  \operatorname{Disc}_s\mathcal M_{\beta\alpha}
  =
  \ii
  \sum_X
  \int \dd\Phi_X\,
  \mathcal M_{X\alpha}\,
  \mathcal M_{X\beta}^{\,*}.
  \label{eq:cutkosky-unitarity-discontinuity}
\end{equation}
Here \(X\) runs over on-shell intermediate states with total momentum
\(P_\alpha\), and \(\dd\Phi_X\) is the Lorentz-invariant phase-space measure,
including the factor \(1/N!\) for \(N\) identical particles when the state sum is
over unordered particles.  The subscript convention is
\(\mathcal M_{\gamma\delta}=\mathcal M(\gamma|\delta)\), so the conjugated
factor is the amplitude \(\beta\to X\), complex conjugated.  It is not, in a
generic theory, the complex conjugate of the reversed process \(X\to\beta\).
Writing \(\mathcal M_{\beta X}^{\,*}\) in this exact Hilbert-space identity
would add a reciprocity assumption, such as time-reversal invariance with
specified antiunitary phase conventions.  The forward case \(\beta=\alpha\)
then reduces to the positive optical-theorem sum.

\begin{remark}[Unitarity discontinuity and perturbative cut rules]
\label{rem:unitarity-discontinuity-versus-cutkosky}
Equation~\eqref{eq:cutkosky-unitarity-discontinuity} is the
nonperturbative on-shell unitarity identity: it relates boundary values of the
exact amplitude to a sum over exact on-shell intermediate states and exact
lower-point amplitudes.  Its hypotheses are Hilbert-space unitarity, the
existence of the relevant scattering states and boundary values, and the
completeness of the state sum used in the channel.  The Cutkosky rule is the
perturbative refinement obtained after choosing a regulator and expansion,
writing \(T=\sum_r T^{(r)}\), representing the terms by Feynman graphs, and
identifying the contribution of a specified graph cut with on-shell
propagator replacements.  Thus the full discontinuity formula and the
graph-by-graph replacement theorem have different logical status, even though
the latter is obtained by expanding the former.
\end{remark}

\begin{proposition}[Largest-time identity for a scalar graph]
\label{prop:largest-time-identity}
Consider a regularized scalar Feynman graph \(G\) with vertices
\(x_1,\ldots,x_V\), local Hermitian interaction vertices, and internal lines
of masses \(m_e\).  For each mass set
\[
  \Delta_m^+(x)
  =
  \int
  \frac{\dd^D q}{(2\pi)^D}\,
  (2\pi)\theta(q^0)\delta(q^2+m^2)\,\ee^{\ii q\cdot x},
  \qquad
  \Delta_m^-(x)=\Delta_m^+(-x),
\]
and decompose the Feynman boundary value as
\[
  \Delta_{F,m}(x)
  =
  \theta(x^0)\Delta_m^+(x)+\theta(-x^0)\Delta_m^-(x),
  \qquad
  \Delta_{F,m}(x)^*
  =
  \theta(x^0)\Delta_m^-(x)+\theta(-x^0)\Delta_m^+(x).
\]
For a circling \(\sigma:\{1,\ldots,V\}\to\{0,1\}\), define \(G_\sigma\) by
using ordinary vertex factors and Feynman propagators inside the uncircled
set, complex-conjugated vertex factors and anti-Feynman propagators inside the
circled set, and Wightman propagators on mixed lines.  A mixed line oriented
from the uncircled side to the circled side carries \(\Delta^+\); the reverse
orientation carries \(\Delta^-\).  Circled vertices carry the minus sign
associated with complex conjugating the perturbative \(i\)-weighted vertex.
Then
\begin{equation}
  \sum_{\sigma\in\{0,1\}^V}G_\sigma(x_1,\ldots,x_V)=0 .
  \label{eq:largest-time-equation}
\end{equation}
\end{proposition}

\begin{proof}
First assume no two vertices have equal time coordinate; the general
distributional statement follows by smearing and partitioning the test
function support away from equal-time hyperplanes, then taking the finite
sum of limits.  Let \(x_k^0\) be the largest vertex time.  Pair the term
\(G_\sigma\) with the term obtained by replacing \(\sigma(k)\) by
\(1-\sigma(k)\) and leaving all other circling labels fixed.  For every line
from \(x_k\) to an earlier vertex, the displayed decompositions give the same
positive-frequency Wightman factor in the two paired terms.  The remaining
factors in the graph are identical.  The two vertex factors at \(x_k\) differ
by the circled-vertex minus sign.  The pair therefore cancels.  Summing over
all circlings proves \eqref{eq:largest-time-equation}.
\end{proof}

Subtract from \eqref{eq:largest-time-equation} the all-uncircled and
all-circled terms.  The former is the original Feynman graph.  The latter is
the complex-conjugated graph with the opposite boundary prescription.  Every
remaining term contains at least one mixed line, and each mixed line has the
Fourier representation
\[
  \frac{\dd^D q}{(2\pi)^D}\,
  (2\pi)\theta(q^0)\delta(q^2+m^2)
\]
after orienting \(q\) from the uncircled side to the circled side.  Thus the
largest-time identity expresses the difference between the two boundary
values of a graph as a sum of cut graphs.  A physical \(s\)-channel cut is the
subsum of those circlings for which the mixed lines separate the external
legs into the chosen left and right subprocesses and carry total momentum
\(P_\alpha\) across the cut.  Circlings whose mixed-line energy support
belongs to another channel contribute to the corresponding channel
discontinuity instead.

\begin{theorem}[Cutkosky cut rule for a perturbative physical cut]
Consider a perturbative contribution to a connected amplitude whose
\(s\)-channel discontinuity is produced by putting a set \(C\) of internal
lines on shell so that the graph separates into a left subamplitude and a right
subamplitude.  With the amplitude normalization fixed by \(S=\mathbf 1+\ii T\),
the discontinuity contribution of this cut is obtained by replacing each cut
propagating line of mass \(m_i\) by the on-shell measure
\[
  \frac{\dd^D q_i}{(2\pi)^D}\,
  (2\pi)\theta(q_i^0)\delta(q_i^2+m_i^2),
\]
multiplying the left and right lower-order amplitudes, complex conjugating the
amplitude on the lower-edge side, and integrating over the common on-shell
phase space.  Equivalently, the graph contribution is the corresponding term in
\eqref{eq:cutkosky-unitarity-discontinuity}.
\end{theorem}

\begin{proof}
There are two statements to match.  The first is the Hilbert-space
discontinuity.  Expand \(T=\sum_r T^{(r)}\) in the chosen interaction order.
The order-\(N\) part of \(-\ii(T-T^\dagger)=T^\dagger T\) is
\[
  -\ii\bigl(T^{(N)}-T^{(N)\dagger}\bigr)
  =
  \sum_{r=1}^{N-1}T^{(r)\dagger}T^{(N-r)} .
\]
Insert the spectral resolution of the identity on the physical Hilbert space
between the two factors on the right.  For a sector of \(n\) stable particles of
masses \(m_i\), this resolution contains
\[
  \frac1{n!}
  \prod_{i=1}^n
  \left[
    \int\frac{\dd^D q_i}{(2\pi)^{D-1}}\,
    \theta(q_i^0)\delta(q_i^2+m_i^2)
  \right]
  (2\pi)^D\delta^{(D)}\!\left(P_\alpha-\sum_i q_i\right).
\]
After the external LSZ residues and the overall momentum-conserving delta
function are removed, this is precisely the phase-space integral in
\eqref{eq:cutkosky-unitarity-discontinuity}.

The second statement identifies the same term inside a fixed Feynman graph.
Apply Proposition~\ref{prop:largest-time-identity} to the graph.  The
difference between its two physical boundary values is the sum of mixed
circlings.  For the chosen \(s\)-channel discontinuity, retain the circlings
whose mixed lines separate the graph into the ordinary subamplitude
\(\alpha\to X\) and the adjoint subamplitude whose unconjugated orientation is
\(\beta\to X\).  The mixed propagator on each separating line is exactly
\[
  \frac{\dd^D q_i}{(2\pi)^D}
  (2\pi)\theta(q_i^0)\delta(q_i^2+m_i^2),
\]
with \(q_i\) oriented from the lower-order amplitude to its conjugate across
the cut.  The uncircled side contains the ordinary Feynman rules for one
lower-order amplitude, while the circled side contains the complex-conjugated
rules for the lower-edge amplitude.  The product of the mixed-line
distributions supplies the on-shell phase-space measure and the residual
momentum delta function enforces \(P_X=P_\alpha\).  This is the graph-level
line-replacement rule stated in the theorem, with the overall factor fixed by
the \(S=\mathbf1+\ii T\) convention used in
\eqref{eq:cutkosky-unitarity-discontinuity}.
\end{proof}

For the two-particle equal-mass cut, the phase-space factor appearing in
\eqref{eq:cutkosky-unitarity-discontinuity} is explicit.  Let \(P^2=-s\) and
\[
  \dd\Phi_2(P)
  =
  \int
  \frac{\dd^4 q}{(2\pi)^3}\theta(q^0)\delta(q^2+m^2)\,
  \frac{\dd^4 r}{(2\pi)^3}\theta(r^0)\delta(r^2+m^2)\,
  (2\pi)^4\delta^{(4)}(P-q-r).
\]
In the center-of-mass frame \(P=(\sqrt s,\vec0)\) this becomes
\[
  \dd\Phi_2(P)
  =
  \frac{\dd^3\vec q}{(2\pi)^2\,4\omega_{\vec q}^{\,2}}\,
  \delta(\sqrt s-2\omega_{\vec q}).
\]
Using \(\omega_{\vec q}=\sqrt{\vec q^{\,2}+m^2}\), the radial delta function
fixes
\[
  |\vec q|=\frac12\sqrt{s-4m^2},
\]
and hence
\begin{equation}
  \int\dd\Phi_2(P)
  =
  \frac1{8\pi}
  \sqrt{1-\frac{4m^2}{s}}\,
  \theta(s-4m^2).
  \label{eq:two-body-phase-space-four-d}
\end{equation}
For identical intermediate particles this is multiplied by \(1/2!\) if the
state sum is over unordered two-particle states.  Thus a one-loop bubble cut
has discontinuity equal to \(\ii\) times the product of the two tree
subamplitudes times \eqref{eq:two-body-phase-space-four-d}, with this symmetry
factor when appropriate.  The square-root branch point is the same threshold
that the Landau equations identify below.

\section{Crossing From Locality and Tube Analyticity}

The structural origin of crossing is the same analytic mechanism used for
Wightman functions.  For four scalar fields, the vacuum distribution
\[
  W_4(x_1,x_2,x_3,x_4)
  =
  \bra{\vac}
  \widehat\phi(x_1)\widehat\phi(x_2)
  \widehat\phi(x_3)\widehat\phi(x_4)
  \ket{\vac}
\]
has Fourier support fixed by the spectrum condition in relative variables.
It is therefore the boundary value of a holomorphic function in a tube
domain.  More explicitly, after removing the center-of-mass coordinate, the
Fourier transform has support in products of closed forward lightcones for
appropriate partial sums of momenta.  This support condition is the input
which makes the imaginary spacetime variables point into forward tubes.

Local commutativity says that, at real configurations where adjacent
insertions are spacelike separated,
\[
  W_4(\ldots,x_i,x_{i+1},\ldots)
  =
  W_4(\ldots,x_{i+1},x_i,\ldots)
\]
for identical scalar fields.  The edge-of-the-wedge theorem then identifies
the analytic continuations attached to the two orderings in the domain
generated from such spacelike configurations.

LSZ reduction is applied to the time-ordered or Wightman boundary values after
isolating the one-particle poles of the external legs.  Different assignments
of external legs as incoming or outgoing particles correspond to different
real boundary components of the same analytically continued four-point
function.  The \(s\)-channel process
\[
  1+2\to3+4
\]
and the crossed process
\[
  1+\bar{3}\to\bar{2}+4
\]
are therefore boundary values of one analytic object in the domain obtained
from permuted Wightman tubes by edge-of-the-wedge continuation.  For a
self-conjugate scalar \(\bar{2}=2\) and \(\bar{3}=3\); for charged fields the
bars indicate the conjugate particle associated with the adjoint field.

The statement is local on the analytic domain.  It does not erase the
singularities separating physical regions.  Threshold cuts, bound-state poles,
and anomalous thresholds specify how the analytic continuation must be routed
and which boundary value is obtained at the end.  Crossing is the assertion of
common analytic origin; the singularity analysis below describes the
obstructions and boundary prescriptions encountered along particular
continuations.

\begin{warning}[BEG crossing versus a Mandelstam domain]
The edge-of-the-wedge argument used here is the rigorous
Bros--Epstein--Glaser mechanism.  It proves that permuted Wightman boundary
values and their LSZ residues have a common analytic continuation on the
envelope generated by the corresponding tube domains.  A path connecting the
\(s\)-channel physical region to a crossed physical region on the same
first-sheet branch, while avoiding all singularities, is additional analytic
data.  That physical-region-to-physical-region crossing statement is therefore
a stronger analyticity assertion.  In perturbation theory it is studied by
tracking graph singularities and sheets; in nonperturbative four-dimensional
QFT it is usually formulated as a Mandelstam-type domain hypothesis rather
than as a theorem following solely from locality, spectrum, and LSZ.
\end{warning}

\section{Edge-of-the-Wedge Form}

The analytic gluing used above has a precise local form.  Let \(F_+\) and
\(F_-\) be holomorphic functions in two domains
\[
  \mathcal D_+,\qquad \mathcal D_-
\]
of the complexified kinematic variables.  Suppose both domains have boundary
values on a common real open set \(U\), in the sense of distributional limits
against test functions supported in \(U\).  Suppose these boundary values
agree:
\[
  F_+|_U=F_-|_U .
\]
The edge-of-the-wedge theorem gives a holomorphic function \(F\) on a domain
containing \(\mathcal D_+\cup U\cup\mathcal D_-\) in its envelope of
holomorphy, with \(F=F_+\) on \(\mathcal D_+\) and \(F=F_-\) on
\(\mathcal D_-\).

In local QFT, the real set \(U\) is supplied by spacelike configurations where
local commutativity equates two Wightman orderings.  The domains
\(\mathcal D_\pm\) are the corresponding forward-tube domains selected by the
spectrum condition.  The theorem then produces a single analytic function
whose different real boundary components give differently ordered
correlators.  LSZ reduction transfers this analytic relation from correlators
to scattering amplitudes whenever the external one-particle pole hypotheses
hold.

\section{Pinches}

On the first sheet one finds stable one-particle poles, ordinary thresholds,
and anomalous thresholds.  In perturbative graph analysis an anomalous
threshold is a contour-pinch singularity on the physical sheet, caused by
several internal lines going on shell in a spacetime arrangement satisfying
the positive Coleman--Norton orientation condition stated in
\eqref{eq:positive-coleman-norton-realization}.  Its invariant location is
determined by the Landau configuration of those internal lines.

The singularities of perturbative amplitudes arise when an integration contour
is forced against singularities of the integrand.  Consider a scalar graph with
internal momenta
\[
  q_i=q_i(\ell_1,\ldots,\ell_L;p_1,\ldots,p_n),
\]
where the \(q_i\) are affine functions of the loop momenta \(\ell_a\) and
external momenta \(p_j\).  A typical denominator is
\[
  q_i^2+m_i^2-\ii0 .
\]
After Feynman parametrization, a product of \(N\) denominators has the form
\[
  \int_{\alpha_i\ge0}
  \prod_{i=1}^N\dd\alpha_i\,
  \delta\!\left(\sum_{i=1}^N\alpha_i-1\right)
  \frac{\text{numerator}}
       {\left[\sum_{i=1}^N\alpha_i(q_i^2+m_i^2)-\ii0\right]^N}.
\]
Numerator factors and ultraviolet subtractions can change the degree or even
remove a candidate singularity, but they do not change the necessary pinch
conditions for the scalar denominator.

As external momenta move in the complex domain, the loop contours can be
deformed as long as no singularity traps them.  A necessary condition for a
singularity is that the denominator vanish at a point where its first
variation in every loop-momentum direction also vanishes.  If the first
variation were nonzero, the contour could be displaced locally away from the
zero.  At a stationary zero, nearby singular hypersurfaces can pinch the
original real contour, and local deformations cannot remove the obstruction.

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.95]
  \draw[blue!70!black,line width=1.25pt,->]
    (-3.3,-0.75) .. controls (-2.2,0.55) and (-0.6,0.95) .. (2.9,0.55);
  \fill[red!75!black] (-0.4,0.68) circle (2.5pt);
  \node[red!75!black,above] at (-0.4,0.78) {pole};
  \fill[orange!90!black] (-0.42,0.43) circle (2.5pt);
  \node[orange!90!black,left] at (-0.55,0.38) {contour};
  \draw[orange!90!black,->,line width=1.0pt] (-0.42,1.12)--(-0.42,0.55);
  \draw[orange!90!black,->,line width=1.0pt] (-0.42,-0.12)--(-0.42,0.30);
  \node[align=center] at (2.45,-0.9)
    {pinch: both local\\deformations are blocked};
\end{tikzpicture}
\end{center}

This stationary-pinch condition is the source of the Landau equations.  The
equations are necessary conditions for a singularity of a graph.  To decide
whether a singularity is actually present one must also check the sheet, the
orientation of the pinch, and whether the numerator cancels the local
singular term.

\section{Landau Equations}

Let
\[
  \mathcal D(\ell,\alpha;p)
  =
  \sum_{i=1}^N\alpha_i(q_i^2+m_i^2).
\]
A Landau singularity can occur only if there are Feynman parameters
\(\alpha_i\ge0\), not all zero, such that
\[
  \alpha_i(q_i^2+m_i^2)=0
  \qquad\text{for every internal line }i,
\]
and, for every loop momentum \(\ell_a^\mu\),
\[
  \frac{\partial\mathcal D}{\partial \ell_a^\mu}=0.
\]
Equivalently,
\[
  \sum_{i=1}^N
  \alpha_i q_{i\nu}
  \frac{\partial q_i^\nu}{\partial \ell_a^\mu}
  =
  0,
  \qquad a=1,\ldots,L .
\]
The omitted factor of \(2\) is irrelevant because the right hand side is
zero.

For a one-loop graph with all internal momenta oriented around the loop, this
reduces to
\[
  \sum_i\alpha_i q_i^\mu=0 .
\]

The first equation says that every line with \(\alpha_i>0\) is on shell:
\[
  q_i^2=-m_i^2 .
\]
The second equation says that the corresponding on-shell momenta form a
stationary configuration around each loop.  Solutions with some
\(\alpha_i=0\) describe singularities of reduced graphs in which the inactive
lines have been contracted.  Solutions with all relevant \(\alpha_i>0\) are
the leading Landau singularities of that graph.

With the terminology fixed in
Section~\ref{sec:physical-region-and-first-sheet}, the physical sheet is
selected by the sign of the Feynman prescription together with positive
Feynman parameters.  Positivity of the Feynman parameters distinguishes a
first-sheet threshold from the same algebraic locus reached only after
continuation through a cut.

\section{Spacetime Interpretation of Positive Landau Solutions}

The sign condition on the parameters has a spacetime meaning.  Consider a
Landau solution in a real physical region and keep only the active internal
lines, namely those with \(\alpha_i>0\).  Let the graph vertices be labelled
by \(v\), and orient each active edge \(i\) from a tail \(v_-(i)\) to a head
\(v_+(i)\).  The momentum \(q_i\) is the momentum assigned to this oriented
edge by the chosen loop-momentum routing.  The loop-stationarity equations say
that, for every oriented cycle \(C\) of the active graph,
\[
  \sum_{i\in C}\sigma_{Ci}\alpha_i q_i=0,\qquad
  \sigma_{Ci}=
  \begin{cases}
    +1,&\text{if the edge orientation agrees with the orientation of }C,\\
    -1,&\text{if it is opposite.}
  \end{cases}
\]
Equivalently, the vector assigned to an active edge by \(\alpha_i q_i\) has
zero sum around every closed path.  Hence, on each connected component, there
are vertex positions \(x_v\), unique up to a common translation, such that
\[
  x_{v_+(i)}-x_{v_-(i)}=\lambda_i q_i,
  \qquad
  \lambda_i\ge0 ,
\]
with \(\lambda_i\) proportional to \(\alpha_i\).  The proportionality constant
depends only on the convention used for the quadratic form in the parametrized
denominator.

The positive-energy condition is an additional orientation condition.  Choose a
time orientation in Minkowski space.  A positive Coleman--Norton realization of
the Landau solution consists of an orientation of every active edge in its
classical direction of propagation and momenta \(k_i\) on those oriented edges
such that
\begin{equation}
  k_i^2=-m_i^2,\qquad k_i^0>0,\qquad
  x_{v_+(i)}-x_{v_-(i)}=\lambda_i k_i,\qquad
  \lambda_i\ge0 .
  \label{eq:positive-coleman-norton-realization}
\end{equation}
If the loop-momentum routing originally assigned the opposite orientation to
that edge, then \(k_i=-q_i\) and the tail and head used in the spacetime
equation are interchanged.  Equivalently, one may reorient the active graph so
that \(q_i=k_i\) on every active edge.  Thus the compatibility condition is
precisely that each active internal momentum be a future-directed
positive-energy on-shell momentum in the direction in which the corresponding
edge is traversed in spacetime.

For \(m_i>0\), this is the worldline equation for a classical massive particle
propagating for nonnegative proper time, with \(\lambda_i m_i\) proportional to
that proper time.  For \(m_i=0\), it is the corresponding future-directed null
geodesic equation with nonnegative affine parameter.  Momentum conservation at
each vertex is the graph momentum conservation for the same oriented momenta,
with the external momenta assigned to that vertex.  Thus a real leading Landau
solution satisfying the positive Coleman--Norton orientation condition is a
classical spacetime process: all active internal particles go on shell, carry
positive energy in their direction of propagation, and meet at the vertices
with energy and momentum conserved.  This is the Coleman--Norton
interpretation.  It is a criterion for first-sheet physical singularities that
must be combined with the pinch analysis: a formal algebraic solution
contributes to the physical singularity only on the relevant sheet, with the
integration contour trapped, and with a numerator that leaves a nonzero local
singular term.

\begin{warning}[Scope of the Coleman--Norton converse]
The implication just proved is the geometric interpretation of a real positive
Landau solution: such a solution is a classical spacetime process.  The
converse statement, often phrased as ``every allowed classical process gives a
singularity,'' has a narrower status.  For a fixed perturbative Feynman graph,
a nondegenerate positive pinch on the relevant sheet produces a local
singularity of the scalar integral; numerator factors, symmetry factors,
subtractions, or sums of graphs may cancel the leading local term.  Thus the
perturbative converse is a theorem only after the pinch orientation,
nondegeneracy, and non-cancellation hypotheses have been checked.  In an exact
nonperturbative QFT there is no general theorem that every such classical
spacetime picture produces a singularity of the full \(S\)-matrix.  There it
should be regarded as a diagnostic expectation, to be justified from the
spectral and locality structure of the theory in each setting.
\end{warning}

Subleading Landau solutions have some \(\alpha_i=0\).  Geometrically, the
corresponding internal lines are contracted, and the classical propagation
picture applies to the reduced graph.  Ordinary thresholds are reduced
classical processes in which a set of particles can be produced on shell with
the total external energy-momentum supplied to the subgraph.

\section{Two-Particle Threshold}

For the equal-mass two-propagator loop, take internal momenta
\[
  q_1=\ell,\qquad q_2=\ell+P .
\]
The Landau equations with \(0<\alpha<1\) are
\[
  \ell^2+m^2=0,\qquad
  (\ell+P)^2+m^2=0,
\]
and
\[
  (1-\alpha)\ell+\alpha(\ell+P)=0 .
\]
The stationary equation gives
\[
  \ell=-\alpha P .
\]
Substitution into the two on-shell equations gives
\[
  \alpha=\frac12,\qquad
  \ell=-\frac{P}{2},\qquad
  P^2=-4m^2 .
\]
Thus the ordinary two-particle threshold is a Landau singularity.
In terms of the invariant \(s=-P^2\), this is \(s=4m^2\).

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.92]
  \begin{scope}
    \draw[line width=0.85pt] (-1.55,0.52)--(-0.48,0.52);
    \draw[line width=0.85pt] (-1.55,-0.52)--(-0.48,-0.52);
    \draw[line width=0.85pt] (0.48,0.52)--(1.55,0.52);
    \draw[line width=0.85pt] (0.48,-0.52)--(1.55,-0.52);
    \draw[line width=0.85pt] (0,0) ellipse (0.58 and 0.5);
    \node[blue!70!black] at (0,0.9) {\(q_2=\ell+P\)};
    \node[blue!70!black] at (0,-0.9) {\(q_1=\ell\)};
    \node at (0,-1.45) {bubble graph};
  \end{scope}
  \begin{scope}[xshift=5.3cm]
    \draw[->,line width=1.0pt] (0,0)--(-1.25,0) node[left] {\(q_1\)};
    \draw[->,line width=1.0pt] (0,0)--(1.25,0) node[right] {\(q_2\)};
    \draw[blue!70!black,->,line width=0.95pt] (1.25,-0.28)--(-1.25,-0.28)
      node[midway,below] {\(P\)};
    \node[align=center] at (0,0.85)
      {stationary point\\\(\alpha=\frac12\)};
    \node[blue!70!black] at (0,-1.05) {\(P^2=-4m^2\)};
  \end{scope}
\end{tikzpicture}
\end{center}

For unequal masses, the same equations give
\[
  \ell=-\alpha P,\qquad
  \alpha^2P^2=-m_1^2,\qquad
  (1-\alpha)^2P^2=-m_2^2 .
\]
The positive-parameter solution with \(0<\alpha<1\) is
\[
  \alpha=\frac{m_1}{m_1+m_2},
  \qquad
  P^2=-(m_1+m_2)^2 .
\]
This is the ordinary threshold: the two on-shell internal particles propagate
with zero relative spatial momentum in the center-of-mass frame.  The
algebraic equations also contain the pseudo-threshold
\[
  P^2=-(m_1-m_2)^2,
\]
but its parameter signs do not describe the same first-sheet physical
production process.

\section{Triangle Singularities}

The triangle graph is the first elementary example of a first-sheet singularity
produced by a genuine on-shell contour pinch rather than by a pole or endpoint
threshold. Let the three external momenta be
\[
  P_1^2=-M_1^2,\qquad
  P_2^2=-M_2^2,\qquad
  P_3^2=-M_3^2 ,
\]
with \(P_1+P_2+P_3=0\). The three internal lines are required to be on shell,
and the positive-parameter stationary condition requires their momentum
vectors to close consistently around the loop.
For one choice of loop orientation this means
\[
  q_i^2=-m_i^2,\qquad
  \alpha_1q_1+\alpha_2q_2+\alpha_3q_3=0,
  \qquad
  \alpha_i>0 .
\]
Together with momentum conservation at the vertices, these equations define
the candidate singular locus in the external invariants.  Equivalently, the
Gram matrix \(G_{ij}=q_i\cdot q_j\) has a null vector
\((\alpha_1,\alpha_2,\alpha_3)\) with strictly positive entries.  Thus the
condition is not only \(\det G=0\), but the existence of a positive null vector
for the oriented on-shell momenta.  The graph and vector-closure data are shown
in Figure~\ref{fig:triangle-landau-positive-parameter-data}.

\begin{figure}[!ht]
\centering
\begin{tikzpicture}[>=Stealth,scale=0.86]
  \begin{scope}
    \coordinate (a) at (-1.25,0.25);
    \coordinate (b) at (1.0,0.78);
    \coordinate (c) at (0.55,-1.05);
    \draw[line width=0.9pt] (a)--(b)--(c)--cycle;
    \foreach \p in {a,b,c}
      \fill[white,draw,line width=0.75pt] (\p) circle (0.15);
    \draw[blue!70!black,->,line width=1pt] (-2.35,0.25)--(a);
    \node[blue!70!black,above] at (-1.95,0.28) {\(P_1\)};
    \draw[blue!70!black,->,line width=1pt] (1.75,1.45)--(b);
    \node[blue!70!black,right] at (1.55,1.18) {\(P_3\)};
    \draw[blue!70!black,->,line width=1pt] (0.68,-2.15)--(c);
    \node[blue!70!black,right] at (0.62,-1.72) {\(P_2\)};
    \node at (0,-2.42) {triangle graph};
  \end{scope}
  \begin{scope}[xshift=5.35cm,yshift=-0.15cm]
    \draw[->,line width=1.0pt] (0,0)--(1.45,0);
    \node[right] at (1.53,0.08) {\(q_1\)};
    \draw[->,line width=1.0pt] (0,0)--(0.82,1.18);
    \node[above] at (0.78,1.35) {\(q_2\)};
    \draw[->,line width=1.0pt] (0,0)--(0.82,-1.18);
    \node[fill=white,inner sep=1pt] at (0.62,-0.74) {\(q_3\)};
    \draw[blue!70!black,->,line width=0.95pt] (0.82,1.18)--(1.82,0.38);
    \node[blue!70!black,right] at (1.9,0.55) {\(M_2\)};
    \draw[blue!70!black,->,line width=0.95pt] (0.82,-1.18)--(1.82,-0.38);
    \node[blue!70!black,right] at (1.9,-0.55) {\(M_1\)};
    \draw[blue!70!black,<->,line width=0.95pt] (0.22,-1.18)--(0.22,1.18);
    \node[blue!70!black,left] at (0.1,0) {\(M_3\)};
    \node[align=center] at (1.28,-2.12)
      {positive-parameter\\vector closure};
  \end{scope}
\end{tikzpicture}
\caption{Triangle Landau data.  The left panel labels the external momentum
flow in the triangle graph.  The right panel shows the positive-parameter
closure condition \(\alpha_1q_1+\alpha_2q_2+\alpha_3q_3=0\) for oriented
on-shell internal momenta.}
\label{fig:triangle-landau-positive-parameter-data}
\end{figure}

In the equal-internal-mass planar example, the Euclideanized internal momentum
vectors in the sketch all have length \(m\).  The external invariant \(M_a\)
is the length of the chord between two such vectors, so
\[
  M_a^2=2m^2(1-\cos\varphi_a),
\]
where \(\varphi_a\) is the angle between the corresponding internal on-shell
vectors.  Thus \(M_a=\sqrt2\,m\) is the right-angle value.  If
\(M_1,M_2<\sqrt2\,m\), the relevant vectors lie in an open half-plane in the
configuration shown, and no positive relation
\(\alpha_1q_1+\alpha_2q_2+\alpha_3q_3=0\) exists.  If
\(M_1,M_2>\sqrt2\,m\), the vectors can be arranged with the origin in their
convex hull; the positive-parameter Landau equation can then be satisfied.
The remaining invariant is fixed by this geometry, giving a singularity at
the corresponding value \(P_3^2=-M_3^2\).  The convex-hull distinction is drawn
in Figure~\ref{fig:triangle-landau-convex-hull-criterion}.

\begin{figure}[!ht]
\centering
\begin{tikzpicture}[>=Stealth,scale=0.84]
  \begin{scope}
    \node[align=center] at (0,1.95)
      {\(M_1,M_2<\sqrt2\,m\)};
    \draw[gray!55,dashed] (0,-1.45)--(0,1.45);
    \fill[gray!10] (0,-1.45)--(1.85,-1.45)--(1.85,1.45)--(0,1.45)--cycle;
    \coordinate (o) at (0,0);
    \coordinate (a) at (1.42,0.08);
    \coordinate (b) at (0.82,1.05);
    \coordinate (c) at (0.92,-0.72);
    \draw[green!50!black,dashed,line width=0.7pt] (a)--(b)--(c)--cycle;
    \draw[->,line width=1.0pt] (o)--(a) node[right] {\(q_1\)};
    \draw[->,line width=1.0pt] (o)--(b) node[above] {\(q_2\)};
    \draw[->,line width=1.0pt] (o)--(c) node[below] {\(q_3\)};
    \fill (o) circle (1.8pt);
    \node[align=center] at (0,-1.85)
      {origin outside convex hull\\no positive closure};
  \end{scope}
  \begin{scope}[xshift=6.4cm]
    \node[align=center] at (0,1.95)
      {\(M_1,M_2>\sqrt2\,m\)};
    \coordinate (o) at (0,0);
    \coordinate (a) at (1.28,0.0);
    \coordinate (b) at (-0.54,1.12);
    \coordinate (c) at (-0.66,-1.02);
    \fill[green!14] (a)--(b)--(c)--cycle;
    \draw[green!50!black,dashed,line width=0.75pt] (a)--(b)--(c)--cycle;
    \draw[->,line width=1.0pt] (o)--(a) node[right] {\(q_1\)};
    \draw[->,line width=1.0pt] (o)--(b) node[above left] {\(q_2\)};
    \draw[->,line width=1.0pt] (o)--(c) node[below left] {\(q_3\)};
    \fill (o) circle (1.8pt);
    \draw[blue!70!black,->,line width=0.9pt]
      (-0.1,0.25) .. controls (0.18,0.55) and (0.45,0.45) .. (0.62,0.2);
    \node[align=center] at (0,-1.85)
      {origin inside convex hull\\positive closure possible};
  \end{scope}
\end{tikzpicture}
\caption{Convex-hull form of the positive-parameter Landau condition in the
equal-internal-mass triangle example.  When the origin lies outside the hull of
the oriented on-shell vectors no positive closure exists; when it lies inside,
the stationary equation can be solved with \(\alpha_i>0\).}
\label{fig:triangle-landau-convex-hull-criterion}
\end{figure}

The vector sketch represents the Coleman--Norton condition.  The actual
singular locus is obtained by solving the three on-shell equations together
with the positive-parameter stationary equation.  A solution gives a
first-sheet anomalous threshold at the determined value of \(P_3^2=-M_3^2\)
only when the active edges can be oriented with momenta \(k_i^0>0\) and
separations \(x_{v_+(i)}-x_{v_-(i)}=\lambda_i k_i\), \(\lambda_i\ge0\).  A
solution requiring reversed time ordering or a negative proper-time segment
lies on a continued sheet, even if it solves the same algebraic Landau
equations.

Such a singularity is an anomalous threshold: a first-sheet singularity of
the amplitude produced by on-shell internal propagation and contour pinching.
Its local data are the masses, momentum routing, and positive Feynman
parameters of the Landau solution.  Ordinary endpoint thresholds are the
reduced-graph case in which the on-shell process begins at the boundary of the
integration domain; the triangle is the first elementary interior-pinch case.
Coleman--Thun singularities in two-dimensional factorized scattering give a
nonperturbative setting in which the same on-shell propagation and contour
pinching mechanism appears.

\section{Use of Landau Analysis}

Landau equations are derived graph by graph. Their role here is diagnostic:
they identify the local analytic mechanisms by which thresholds and more
subtle first-sheet singularities arise in perturbation theory. In an exact
massive theory, the singularities of amplitudes are expected to be governed by
the same spectral and locality principles, with internal on-shell lines
replaced by actual stable particles or by spectral continua of the theory.

Within existing axiomatic frameworks, only parts of this analytic picture are
presently theorems.  The terminology used here is precise, but its logical
status is not uniform: stable particles appear as first-sheet poles below
thresholds under the one-particle spectral hypotheses, resonances are
continued-sheet pole data, ordinary thresholds are tied to multi-particle
spectral continua, and anomalous thresholds are perturbative on-shell
propagation patterns encoded by the Landau equations.  Beyond perturbation
theory the last statement is a conjectural guide unless a separate
nonperturbative analytic construction is supplied.  This classification will
be used in the following chapter only at the level needed for angular
analyticity: the nearest first-sheet crossed-channel singularity fixes the
distance from the physical angular interval to the edge of the Lehmann ellipse.
Landau analysis supplies a concrete perturbative mechanism for such
singularities.  The nonperturbative construction of the \(S\)-operator and the
spectral assumptions stated earlier remain the structural input for the exact
scattering problem.
