\chapter{Composite Bound States and Bethe--Salpeter Amplitudes}
\label{ch:volume-ii-composite-bound-states-bethe-salpeter}
\ConventionsLorentzian

The time-ordered four-point functions, residues, and perturbative bubble
integrals in this chapter are understood with the status convention of
Definition~\ref{def:scattering-time-ordered-correlator-status}.  Exact
bound-state pole statements use Hilbert-space spectral and LSZ data; the
bubble-chain calculation is a formal regulated graph calculation of the
corresponding time-ordered kernel.


\section{Composite Particles as Poles}

Let \(\phi\) be a real scalar field whose stable one-particle states have
mass \(m\). A stable two-particle composite species \(B\) of mass \(M_B\) is an
isolated one-particle sector whose wavepacket states are normalizable and whose
mass satisfies
\[
  0<M_B<2m .
\]
The inequality places the state below the two-\(\phi\) continuum threshold.

\begin{assumption}[Bound-state pole hypothesis]
The existence of \(B\) is an input in this chapter.  More explicitly, assume
that the Hilbert space contains an isolated positive-energy one-particle
subrepresentation of mass \(M_B\), separated from the two-\(\phi\) continuum,
and that at least one operator in the two-\(\phi\) channel has nonzero matrix
element with this state.  The following sections derive the consequences of
that spectral hypothesis for four-point functions, scattering amplitudes, and
Bethe--Salpeter residues.  They do not give a general proof that an arbitrary
Lagrangian with attractive interactions has such a state.
\end{assumption}

It is then represented, on the first sheet in the sense of
Section~\ref{sec:physical-region-and-first-sheet}, by a pole of the appropriate
four-point function at
\[
  s=M_B^2,
  \qquad
  s=-(k_1+k_2)^2 .
\]
Equivalently, for the \(2\to2\) invariant amplitude \(\mathcal M(s,t)\), whose
connected \(S\)-kernel transition part is \(\ii\mathcal M(s,t)\) after the
overall momentum-conservation delta function is removed, with
\[
  t=-(k_1-k_3)^2 ,
  \qquad
  u=-(k_1-k_4)^2 ,
  \qquad
  s+t+u=4m^2 ,
\]
the same state appears as a pole in \(s\), at fixed quantum numbers in the
\((12)\)-channel. After LSZ reduction of the external stable \(\phi\)
particles, the pole is present in the connected scattering amplitude; the
external one-particle residues multiply the amplitude and do not determine the
position of the pole.
These are the mostly-plus Mandelstam variables for on-shell external momenta:
\(k_i^2=-m^2\), with \(k_1,k_2\) incoming and \(k_3,k_4\) outgoing.

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.95]
  \begin{scope}
    \draw[->,line width=0.75pt] (-0.35,0)--(6.7,0) node[right] {\(s\)};
    \draw[->,line width=0.75pt] (0,-1.35)--(0,1.9)
      node[above] {\(\operatorname{Im}s\)};
    \fill[blue!75!black] (2.15,0) circle (2.5pt);
    \node[blue!75!black,below] at (2.15,-0.18) {\(M_B^2\)};
    \draw[red!75!black,line width=1.8pt] (4.3,0)--(6.45,0);
    \node[red!75!black,below] at (4.3,-0.18) {\(4m^2\)};
    \node[blue!75!black,align=center] at (2.0,1.0)
      {first-sheet\\composite pole};
    \draw[blue!75!black,->,line width=0.8pt] (2.0,0.68)--(2.12,0.12);
    \node[red!75!black,align=center] at (5.45,0.72)
      {two-particle\\cut};
  \end{scope}
  \begin{scope}[xshift=9.1cm]
    \node[draw,circle,fill=gray!15,minimum size=0.92cm,inner sep=0pt] (b) at (0,0) {};
    \draw[line width=0.85pt] (-1.35,0.82)--(b.145);
    \draw[line width=0.85pt] (-1.35,-0.82)--(b.215);
    \draw[line width=0.85pt] (b.35)--(1.35,0.82);
    \draw[line width=0.85pt] (b.325)--(1.35,-0.82);
    \node at (0,0) {\(\mathcal M\)};
    \node[left] at (-1.35,0.82) {\(k_1\)};
    \node[left] at (-1.35,-0.82) {\(k_2\)};
    \node[right] at (1.35,0.82) {\(k_3\)};
    \node[right] at (1.35,-0.82) {\(k_4\)};
  \end{scope}
\end{tikzpicture}
\end{center}

This statement is independent of whether the Lagrangian contains a field
named \(B\). What matters is the spectral state and the residue with which
operators in the two-particle channel create it.

\section{A Bubble-Chain Example}

The explicit calculation in this section is a \(D=2\) spacetime-dimensional
model.  The purpose is to exhibit, in a finite elementary integral, how a
geometric sum of connected four-point kernels can generate a stable pole below
the two-particle threshold.  In \(D=4\) the same channel has a different
threshold behavior and the pole condition involves the renormalized contact
interaction at threshold; the numerical estimate for \(M_B\) below is therefore
not a dimension-independent formula.
The word generate in this paragraph refers to the specified bubble-chain
approximation.  A zero of the resummed denominator is a pole of that
approximate kernel.  Promoting it to an exact particle statement for a complete
model requires additional nonperturbative control of the spectrum and of the
analytic continuation of the exact four-point function.

Consider the scalar theory
\[
  \mathcal L
  =
  -\frac12(\partial_\mu\phi)^2
  -\frac12m^2\phi^2
  -\frac{g}{4!}\phi^4 .
\]
The four-point vertex is \(-\ii g\).  Before specializing the integral, the
\(s\)-channel bubble-chain approximation to \(2\to2\) scattering contains the
one-loop bubble
\[
  \frac{(-\ii g)^2}{2}
  \int\frac{\dd^D\ell}{(2\pi)^D}\,
  \frac{-\ii}{\ell^2+m^2-\ii0}\,
  \frac{-\ii}{(k_1+k_2-\ell)^2+m^2-\ii0}.
\]
The factor \(1/2\) is the symmetry factor of the two identical internal
lines.  Feynman parametrization and Wick rotation give
\[
  \frac{\ii g^2}{2}
  \int_0^1\dd x
  \int\frac{\dd^D\ell_E}{(2\pi)^D}\,
  \frac{1}{\left[\ell_E^2-sx(1-x)+m^2\right]^2}.
\]
As in the self-energy calculation, the Wick-rotated denominator is real on the
Euclidean side; the \(-\ii0\) prescription enters only when the resulting
function is continued back to a Lorentzian boundary value.

Specializing now to \(D=2\), this becomes
\[
  \frac{\ii g^2}{8\pi} f(s),
  \qquad
  f(s)=
  \int_0^1\dd x\,
  \frac{1}{m^2-sx(1-x)-\ii0}.
\]
With the convention of
Chapter~\ref{ch:volume-ii-scattering-kernels}, the geometric bubble chain gives
the connected \(S\)-kernel transition part
\[
  \ii\mathcal M_{\mathrm{chain}}(s)
  =
  -\ii g
  \sum_{L=0}^{\infty}
  \left(-\frac{g}{8\pi}f(s)\right)^L
  =
  \frac{-\ii g}{1+\dfrac{g}{8\pi}f(s)} .
\]
External wavefunction factors multiply the expression and do not move the
zeros of the denominator.

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.82,
  leg/.style={line width=0.8pt},
  dot/.style={fill=gray!25,draw,circle,inner sep=1.8pt}]
  \begin{scope}
    \node at (-0.9,0) {\(\ii\mathcal M_{\mathrm{chain}}=\)};
    \node[dot] (c) at (0.35,0) {};
    \draw[leg] (-0.35,0.55)--(c)--(1.05,0.55);
    \draw[leg] (-0.35,-0.55)--(c)--(1.05,-0.55);
    \node at (1.55,0) {\(+\)};
    \begin{scope}[xshift=2.6cm]
      \draw[leg] (-0.75,0.55)--(-0.2,0.16);
      \draw[leg] (-0.75,-0.55)--(-0.2,-0.16);
      \draw[leg] (0.2,0.16)--(0.75,0.55);
      \draw[leg] (0.2,-0.16)--(0.75,-0.55);
      \draw[leg] (0,0) circle (0.34);
      \node at (1.12,0) {\(+\)};
    \end{scope}
    \begin{scope}[xshift=4.9cm]
      \draw[leg] (-0.95,0.55)--(-0.48,0.16);
      \draw[leg] (-0.95,-0.55)--(-0.48,-0.16);
      \draw[leg] (0.48,0.16)--(0.95,0.55);
      \draw[leg] (0.48,-0.16)--(0.95,-0.55);
      \draw[leg] (-0.26,0) circle (0.30);
      \draw[leg] (0.26,0) circle (0.30);
      \node at (1.38,0) {\(+\cdots\)};
    \end{scope}
    \node[align=center,blue!70!black,font=\small] at (3.25,-1.25)
      {pole condition:\\\(1+g f(s)/(8\pi)=0\)};
  \end{scope}
\end{tikzpicture}
\end{center}

Near the two-particle threshold from below, write
\[
  s=4m^2-\delta,
  \qquad
  \delta\to0^+ .
\]
The integral is dominated by \(x=1/2\). Setting \(x=1/2+y\), one obtains
\[
  m^2-sx(1-x)
  =
  \frac{\delta}{4}+4m^2y^2+O(\delta y^2),
\]
and hence
\[
  f(4m^2-\delta)
  \simeq
  \frac{\pi}{m\sqrt{\delta}} .
\]
The denominator can vanish for attractive coupling \(g<0\). At weak
coupling,
\[
  \delta
  \simeq
  \left(\frac{g}{8m}\right)^2,
  \qquad
  M_B
  =
  \sqrt{4m^2-\delta}
  \simeq
  2m-\frac{g^2}{256m^3}.
\]
Thus, within the \(D=2\) bubble-chain approximation and for attractive weak
coupling, the resummed denominator has a stable-pole zero slightly below the
two-particle threshold.

The calculation uses the local expansion around a perturbative vacuum. A
complete scalar model with an attractive quartic term must specify the
large-field behavior of the potential. For example,
\[
  V(\phi)=\frac{m^2}{\beta^2}\left(1-\cos\beta\phi\right)
  =
  \frac12m^2\phi^2
  -\frac{m^2\beta^2}{4!}\phi^4
  +O(\phi^6)
\]
has an attractive quartic term in its local expansion, with local quartic
coefficient \(g_{\mathrm{local}}=-m^2\beta^2\), while retaining a potential
bounded from below. The bubble-chain computation should therefore be read as a
statement about the local interaction coefficients around the chosen vacuum,
not as a nonperturbative definition of an unstable negative-quartic polynomial
theory.

\section{Bethe--Salpeter Amplitude}

Let \(\ket{B(P)}\) be a stable composite one-particle state of mass \(M_B\),
with
\[
  P^2=-M_B^2,\qquad P^0>0 .
\]
We use the covariant plane-wave normalization
\[
  \langle B(\vec P\,')|B(\vec P)\rangle
  =
  (2\pi)^{D-1}2\omega_{\vec P}\,
  \delta^{(D-1)}(\vec P-\vec P\,'),
  \qquad
  \omega_{\vec P}=\sqrt{\vec P^{\,2}+M_B^2}.
\]
The momentum eigenstate is distributional; physical Hilbert-space states are
obtained by smearing it with square-integrable wavepackets.
The Bethe--Salpeter amplitude associated with the scalar local field
\(\widehat\phi\) is the time-ordered matrix element
\[
  \Phi_B(x_1,x_2;P)
  =
  \bra{\vac}T\widehat\phi(x_1)\widehat\phi(x_2)|B(P)\rangle .
\]
Here \(\Phi_B\) names a c-number bound-state amplitude, not a field operator.
Translation covariance gives
\[
  \Phi_B(x_1+a,x_2+a;P)
  =
  \ee^{\ii P\cdot a}\Phi_B(x_1,x_2;P),
\]
with the Fourier convention used below. Therefore its Fourier transform has
support on total momentum \(P\):
\[
  \Phi_B(x_1,x_2;P)
  =
  \int\frac{\dd^D k_1}{(2\pi)^D}
       \frac{\dd^D k_2}{(2\pi)^D}\,
  \ee^{\ii k_1\cdot x_1+\ii k_2\cdot x_2}
  (2\pi)^D\delta^{(D)}(k_1+k_2-P)
  \widetilde\Phi_B(k_1,k_2;P).
\]
If instead one uses unit-normalized momentum kets, a factor
\(\bigl[(2\pi)^{D-1}2\omega_{\vec P}\bigr]^{-1/2}\) appears on the right-hand
side; here that factor is absorbed into the definition of
\(\widetilde\Phi_B\). The invariant normalization of
\(\widetilde\Phi_B\) is fixed by the four-point pole residue.
No constituent LSZ factors are divided out in this definition.  If the scalar
field has one-particle residue
\[
  \bra{\vac}\widehat\phi(0)|\vec p\rangle=Z_\phi^{1/2},
\]
then \(\widetilde\Phi_B\) is the unamputated, field-normalized
Bethe--Salpeter amplitude and contains the \(Z_\phi^{1/2}\) normalization
associated with each constituent field.  A convention using LSZ-normalized
constituent fields would instead use
\[
  \widetilde\Phi_B^{\rm LSZ}
  =
  Z_\phi^{-1}\widetilde\Phi_B
\]
for two identical scalar constituents.  In this chapter the inverse LSZ factors
\(Z_\phi^{-1/2}\) are not part of \(\widetilde\Phi_B\); they are inserted only
when a four-point Green function is converted to an invariant scattering
amplitude.

Now consider the momentum-space time-ordered four-point function
\[
\begin{aligned}
  \widetilde G_4(k_1,k_2,k_3,k_4)
  &=
  \int\prod_{j=1}^4\dd^D x_j\,
  \exp\!\left[-\ii\sum_{j=1}^4 k_j\cdot x_j\right]
\\
  &\qquad\qquad\times
  \bra{\vac}T\phi(x_1)\phi(x_2)\phi(x_3)\phi(x_4)\ket{\vac} .
\end{aligned}
\]
All four momenta are taken incoming, so translation invariance supplies
\((2\pi)^D\delta^{(D)}(k_1+k_2+k_3+k_4)\).

\begin{theorem}[Bethe--Salpeter pole factorization]
\label{thm:bethe-salpeter-pole-factorization}
Assume that the \((12)\)-channel contains an isolated stable scalar bound
state \(B\) of mass \(M_B\), separated from the multiparticle continuum, and
that after smearing the relative coordinates of the pairs \((12)\) and
\((34)\), the resulting pair operators admit a channel spectral
decomposition with this isolated one-particle subspace.  With
\[
  P=k_1+k_2,\qquad P^2\to -M_B^2,
\]
the pole part of the Fourier-transformed four-point function is
\begin{equation}
\begin{aligned}
  \widetilde G_4(k_1,k_2,k_3,k_4)
  &\sim
  (2\pi)^D\delta^{(D)}(k_1+k_2+k_3+k_4)
\\
  &\quad\times
  \frac{-\ii\,
    \widetilde\Phi_B(k_1,k_2;P)\,
    \widetilde\Phi_B^*(-k_3,-k_4;P)}
       {P^2+M_B^2-\ii0}
  +\text{terms regular at }P^2=-M_B^2 .
\end{aligned}
  \label{eq:bethe-salpeter-four-point-pole-factorization}
\end{equation}
\end{theorem}

\begin{proof}
First prove the pole statement after smearing the relative coordinates.  Let
\(\xi=x_1-x_2\), \(\eta=x_3-x_4\), \(X=(x_1+x_2)/2\), and
\(Y=(x_3+x_4)/2\).  For test functions \(f,g\) in the relative variables set
\[
  \mathcal O_f(X)
  =
  \int\dd^D\xi\,f(\xi)\,
  T\{\phi(X+\xi/2)\phi(X-\xi/2)\},
\]
and define \(\mathcal O_g(Y)\) similarly for the \((34)\) pair.  The
singularity in the total momentum \(P\) is controlled by the Fourier transform
in the cluster separation \(R=X-Y\).  Terms in which the two pairs are not
separated by arbitrarily large positive or negative \(R^0\) have compact
support, or rapid decrease after the relative smearing, in the \(R^0\)
direction and therefore give functions regular at \(P^2=-M_B^2\).  The pole
comes from the two cluster time orderings
\[
  \theta(R^0)\bra{\vac}\mathcal O_f(X)\mathcal O_g(Y)\ket{\vac}
  +
  \theta(-R^0)\bra{\vac}\mathcal O_g(Y)\mathcal O_f(X)\ket{\vac}.
\]

Insert the spectral resolution between the two pair operators.  The isolated
one-particle bound-state subspace contributes
\[
  \int
  \frac{\dd^{D-1}\vec Q}{(2\pi)^{D-1}2\omega_{\vec Q}}\,
  |B(\vec Q)\rangle\langle B(\vec Q)|,
  \qquad
  \omega_{\vec Q}=\sqrt{\vec Q^{\,2}+M_B^2},
\]
with the invariant one-particle normalization used throughout the scattering
chapters.  Translation covariance gives
\[
  \bra{\vac}\mathcal O_f(X)|B(\vec Q)\rangle
  =
  \ee^{\ii Q\cdot X}
  F_f(\vec Q),
  \qquad
  \langle B(\vec Q)|\mathcal O_g(Y)|\vac\rangle
  =
  \ee^{-\ii Q\cdot Y}
  F_g(\vec Q)^*,
\]
where \(Q^0=\omega_{\vec Q}\), and \(F_f,F_g\) are the relative-coordinate
smearings of the Bethe--Salpeter amplitudes.  The bound-state part of the
cluster-ordered correlator is therefore
\[
  \int
  \frac{\dd^{D-1}\vec Q}{(2\pi)^{D-1}2\omega_{\vec Q}}\,
  F_f(\vec Q)F_g(\vec Q)^*
  \left[
    \theta(R^0)\ee^{\ii Q\cdot R}
    +
    \theta(-R^0)\ee^{-\ii Q\cdot R}
  \right],
\]
with the second term obtained from the reversed ordering.  Fourier transform
the cluster separation with total momentum \(P\).  The time integral uses
\[
  \int_{-\infty}^{\infty}\dd\tau\,\ee^{\ii E\tau}
  \left[
    \theta(\tau)\ee^{-\ii\omega\tau}
    +
    \theta(-\tau)\ee^{\ii\omega\tau}
  \right]
  =
  \frac{\ii}{E-\omega+\ii0}
  -
  \frac{\ii}{E+\omega-\ii0}
  =
  \frac{2\ii\omega}{E^2-\omega^2+\ii0}.
  \label{eq:bs-time-ordering-feynman-pole}
\]
Combining this with the invariant measure \(1/(2\omega_{\vec Q})\) gives
\[
  \frac{\ii}{(P^0)^2-\omega_{\vec P}^{\,2}+\ii0}
  =
  \frac{-\ii}{P^2+M_B^2-\ii0}
\]
in the mostly-plus metric.  This is the origin of the displayed Feynman
prescription.  The spatial Fourier transform fixes \(\vec Q=\vec P\), and
the remaining common translation gives
\((2\pi)^D\delta^{(D)}(k_1+k_2+k_3+k_4)\).

Removing the relative-coordinate test functions gives the distributional
kernel statement.  The two matrix elements \(F_f\) and \(F_g^*\) unsmear to
\(\widetilde\Phi_B(k_1,k_2;P)\) and
\(\widetilde\Phi_B^*(-k_3,-k_4;P)\).  The continuum spectral integral begins
at invariant mass above \(M_B\), so its denominator has no singularity at
\(P^2=-M_B^2\) and contributes the terms regular at the pole.
\end{proof}

Equation~\eqref{eq:bethe-salpeter-four-point-pole-factorization} is the
four-point analogue of the spectral pole in a two-point function.  The residue
is the product of the amplitudes for two fields to create and annihilate the
stable composite state.

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.9,
  leg/.style={line width=0.85pt},
  blob/.style={draw,fill=gray!15,circle,minimum size=0.92cm,inner sep=0pt},
  bs/.style={draw,fill=green!10,ellipse,minimum width=1.15cm,minimum height=1.0cm,inner sep=0pt}]
  \begin{scope}
    \node[blob] (g) at (0,0) {\(\widetilde G_4\)};
    \draw[leg] (-1.55,0.82)--(g.145);
    \draw[leg] (-1.55,-0.82)--(g.215);
    \draw[leg] (g.35)--(1.55,0.82);
    \draw[leg] (g.325)--(1.55,-0.82);
    \node[left] at (-1.55,0.82) {\(k_1\)};
    \node[left] at (-1.55,-0.82) {\(k_2\)};
    \node[right] at (1.55,0.82) {\(-k_3\)};
    \node[right] at (1.55,-0.82) {\(-k_4\)};
  \end{scope}
  \node at (3.0,0) {\(\sim\)};
  \begin{scope}[xshift=6.0cm]
    \node[bs] (l) at (-1.35,0) {\(\widetilde\Phi_B\)};
    \node[bs] (r) at (1.35,0) {\(\widetilde\Phi_B^*\)};
    \draw[leg] (-3.0,0.82)--(l.145);
    \draw[leg] (-3.0,-0.82)--(l.215);
    \draw[leg] (r.35)--(3.0,0.82);
    \draw[leg] (r.325)--(3.0,-0.82);
    \draw[blue!70!black,line width=1.5pt] (l.east)--(r.west);
    \node[blue!70!black,above] at (0,0.18)
      {\(\displaystyle \frac{-\ii}{P^2+M_B^2-\ii0}\)};
    \node[blue!70!black,below] at (0,-0.28) {\(B\)};
  \end{scope}
\end{tikzpicture}
\end{center}

\section{The Homogeneous Equation}
\label{sec:bethe-salpeter-homogeneous-equation}

Let \(p\) and \(q\) denote relative momenta at fixed total momentum \(P\).
For equal constituent masses one may write the two internal momenta as
\[
  \frac{P}{2}+q,\qquad \frac{P}{2}-q .
\]
Let \(D(k)\) be the exact time-ordered propagator of the constituent field,
and define the two-particle propagator
\[
  \mathcal G_P(q)
  =
  D\!\left(\frac{P}{2}+q\right)
  D\!\left(\frac{P}{2}-q\right).
\]
Let \(\mathcal K_P(p,q)\) be the amputated two-particle irreducible kernel in
the chosen channel: by definition, it cannot be separated into two pieces by
cutting the two constituent lines carrying total momentum \(P\). Contact
vertices, crossed exchanges, and self-energy insertions on a constituent line
may be part of this kernel after the constituent propagator convention has been
fixed; an iteration of two kernels connected by the two constituent lines is
not part of the kernel, because that iteration is precisely what the
Bethe--Salpeter equation resums.

This definition is an operator statement once the relative-momentum variables
are smeared.  Let \(\mathcal X\) be a chosen test-function or Hilbert space of
amputated relative wavefunctions, for instance a weighted Sobolev space on a
Euclidean relative-momentum contour in a massive model.  The distribution
\(\mathcal K_P(p,q)\) represents a kernel only if it induces a continuous map
from \(\mathcal X\) to its dual, or, after the factors of
\(\mathcal G_P\) are included, a densely defined operator on \(\mathcal X\).
Equivalently, in a regulated theory one may define \(\mathcal K_P\) by the
exact four-point block \(\mathcal T_P\) through
\[
  \mathcal K_P
  =
  \mathcal T_P(\mathbf 1+\mathcal G_P\mathcal T_P)^{-1},
\]
on any open \(P^2\)-domain where the inverse exists on the chosen space.  The
continuum kernel is then the limit of this regulated kernel in the same
operator topology.  The present chapter assumes this analytic operator
realization when it writes \(\mathcal K_P\mathcal G_P\) as an integral
operator.  Without such a realization the displayed Bethe--Salpeter equation is
only distributional notation for a desired decomposition of the four-point
function.

The amputated four-point scattering block \(\mathcal T_P(p,p')\) obeys the
operator equation
\begin{equation}
  \mathcal T_P
  =
  \mathcal K_P
  +
  \mathcal K_P\mathcal G_P\mathcal T_P ,
  \label{eq:bethe-salpeter-operator-equation}
\end{equation}
or explicitly
\begin{equation}
\begin{aligned}
  \mathcal T_P(p,p')
  &=
  \mathcal K_P(p,p')
\\
  &\quad+
  \int\frac{\dd^D q}{(2\pi)^D}\,
  \mathcal K_P(p,q)\mathcal G_P(q)\mathcal T_P(q,p') .
\end{aligned}
  \label{eq:bethe-salpeter-integral-equation-for-t}
\end{equation}
A bound-state pole at \(P^2=-M_B^2\) has the form
\begin{equation}
  \mathcal T_P(p,p')
  \sim
  \frac{\Psi_B(p;P)\,\overline{\Psi}_B(p';P)}
       {P^2+M_B^2-\ii0}
  +\text{regular terms}.
  \label{eq:bethe-salpeter-amputated-bound-state-pole}
\end{equation}
Here \(\overline{\Psi}_B\) denotes the residue on the right side of the
four-point function; for scalar constituents it is the complex conjugate with
the momenta reversed, up to the same convention choices already present in
\(\widetilde\Phi_B^*(-k_3,-k_4;P)\).
Substituting this pole form into the integral equation and equating residues
gives the homogeneous Bethe--Salpeter equation
\begin{equation}
  \Psi_B(p;P)
  =
  \int\frac{\dd^D q}{(2\pi)^D}\,
  \mathcal K_P(p,q)\mathcal G_P(q)\Psi_B(q;P),
  \qquad
  P^2=-M_B^2 .
  \label{eq:homogeneous-bethe-salpeter-equation}
\end{equation}
Equivalently, define the integral operator
\[
  (\widehat V_P F)(p)
  =
  \int\frac{\dd^D q}{(2\pi)^D}\,
  \mathcal K_P(p,q)\mathcal G_P(q)F(q).
\]
Then a bound state is a nonzero vector \(\Psi_B\) in the domain of this
operator for which
\[
  (\mathbf 1-\widehat V_P)\Psi_B=0 .
\]
The pole condition is thus an eigenvalue-one condition for the repeated
two-particle propagation at \(P^2=-M_B^2\).
This statement has the logical status of a residue equation unless the
operator problem is supplied with enough analytic structure.  The following
theorem records the standard Fredholm form of the missing structure.  It is
not a general existence theorem for bound states in arbitrary Lagrangians; it
is the precise local theorem that converts an eigenvalue-one Bethe--Salpeter
condition into a pole of the represented four-point block, and conversely,
when the exact kernel has been realized as an analytic compact-operator
family.

\begin{theorem}[Analytic Fredholm Bethe--Salpeter pole criterion]
\label{thm:bethe-salpeter-fredholm-pole-criterion}
Let \(z=P^2\), and let \(U\subset\mathbb C\) be a connected open neighborhood
of \(z_B=-M_B^2\) whose closure is separated from the two-constituent
threshold and from every other singular locus of the chosen first-sheet
continuation.  Let \(\mathcal X\) be the Banach space of amputated
relative-momentum test functions used to realize the exact two-particle
irreducible kernel.  Assume that
\[
  \widehat V(z)=\mathcal K_z\mathcal G_z:\mathcal X\to\mathcal X
\]
is analytic in \(z\) as a family of compact operators, and that
\(\mathbf 1-\widehat V(z_0)\) is invertible for at least one \(z_0\in U\).
Assume also that the part of the amputated four-point block seen by the
external two-field channel has an operator representation
\begin{equation}
  \mathcal T_z^{\mathrm{rep}}
  =
  \mathcal R_z+\mathcal F_z
  \bigl(\mathbf 1-\widehat V(z)\bigr)^{-1}\mathcal E_z ,
  \label{eq:bethe-salpeter-fredholm-representation}
\end{equation}
where \(\mathcal E_z:\mathcal Y\to\mathcal X\),
\(\mathcal F_z:\mathcal X\to\mathcal Y'\), and
\(\mathcal R_z:\mathcal Y\to\mathcal Y'\) are analytic bounded-operator
families for the selected external test space \(\mathcal Y\).

Then \((\mathbf 1-\widehat V(z))^{-1}\), initially defined where the inverse
exists, extends to a meromorphic family on \(U\), and each coefficient in its
principal part has finite rank.  Therefore
\(\mathcal T_z^{\mathrm{rep}}\) is meromorphic on \(U\).  It has a pole at
\(z_B\) precisely when the finite-rank principal part of
\((\mathbf 1-\widehat V(z))^{-1}\) at \(z_B\) is not annihilated by the channel
maps \(\mathcal F_z\) and \(\mathcal E_z\).  In particular, if
\(\ker(\mathbf 1-\widehat V(z_B))\ne0\) and the corresponding root subspace
has nonzero external overlap, then \(z_B\) is a pole of the represented
four-point block.  Conversely, any pole of \(\mathcal T_z^{\mathrm{rep}}\)
whose singular part is represented through
\eqref{eq:bethe-salpeter-fredholm-representation} comes from a nonzero
root vector of \(\mathbf 1-\widehat V(z_B)\).

If the root space is simple, so that
\(\ker(\mathbf 1-\widehat V(z_B))=\mathbb C\psi\),
\(\ker(\mathbf 1-\widehat V(z_B))^*=\mathbb C\chi\),
\(\chi(\psi)=1\), and the corresponding analytic eigenvalue
\(\lambda(z)\) of \(\widehat V(z)\) satisfies
\(\lambda(z_B)=1\), \(\lambda'(z_B)\ne0\), then the singular part is
\begin{equation}
  \mathcal T_z^{\mathrm{rep}}
  =
  -\,\frac{(\mathcal F_{z_B}\psi)\otimes(\chi\mathcal E_{z_B})}
      {\lambda'(z_B)(z-z_B)}
  +\text{\rm holomorphic terms}.
  \label{eq:bethe-salpeter-simple-fredholm-residue}
\end{equation}
The pole is visible in the chosen channel if and only if
\(\mathcal F_{z_B}\psi\ne0\) and \(\chi\mathcal E_{z_B}\ne0\).
When, in addition, \(z_B\) lies on the first sheet below threshold and the
exact Hilbert-space spectral representation has positive one-particle residue
there, this pole is the scattering signature of a stable bound state.
\end{theorem}

\begin{proof}
The analytic Fredholm theorem applies because \(\widehat V(z)\) is analytic
and compact and because \(\mathbf 1-\widehat V(z_0)\) is invertible at one
point of the connected domain \(U\).  Hence
\((\mathbf 1-\widehat V(z))^{-1}\) is meromorphic on \(U\), with singular
coefficients of finite rank supported on the finite-dimensional generalized
root spaces of \(\widehat V(z_B)\) at eigenvalue \(1\).  Multiplication on the
left and right by analytic bounded families \(\mathcal F_z\) and
\(\mathcal E_z\), and addition of the analytic background \(\mathcal R_z\),
preserve meromorphicity and can only remove a singularity by annihilating its
finite-rank principal part.  This proves the pole criterion in
\eqref{eq:bethe-salpeter-fredholm-representation}.

For the simple-root formula, choose analytic right and left eigenvectors
\(\psi(z)\), \(\chi(z)\) near \(z_B\) with
\(\chi(z)\psi(z)=1\).  On the rank-one spectral subspace,
\(\mathbf 1-\widehat V(z)\) acts by multiplication by \(1-\lambda(z)\).
Since
\[
  1-\lambda(z)=-\lambda'(z_B)(z-z_B)+O\bigl((z-z_B)^2\bigr),
\]
the inverse has principal part
\[
  -\,\frac{\psi\otimes\chi}{\lambda'(z_B)(z-z_B)} .
\]
Substitution into \eqref{eq:bethe-salpeter-fredholm-representation} gives
\eqref{eq:bethe-salpeter-simple-fredholm-residue}.  The final sentence is not
a Fredholm consequence: it is the separate spectral identification supplied by
the exact Hilbert-space theory, namely a first-sheet isolated pole below the
continuum with positive residue in a physical channel.
\end{proof}

\begin{remark}[What the homogeneous equation proves]
The homogeneous Bethe--Salpeter equation is therefore an existence theorem
only relative to the hypotheses of
Theorem~\ref{thm:bethe-salpeter-fredholm-pole-criterion}: exact kernel,
specified Banach space, compact analytic dependence on \(P^2\), invertibility
somewhere in the analytic neighborhood, nonzero external overlap, and
spectral positivity for the first-sheet below-threshold pole.  In a ladder or
other truncated kernel, a nonzero solution of
\((\mathbf 1-\widehat V_P)\Psi=0\) is a self-consistency condition inside that
approximation.  It is not, by itself, a derivation of a normalizable
positive-energy bound state in the exact QFT.
\end{remark}

The unamputated Bethe--Salpeter amplitude is related to the amputated vertex by
\[
  \widetilde\Phi_B(p;P)
  =
  \mathcal G_P(p)\Psi_B(p;P),
\]
with the same field normalization used in
\eqref{eq:bethe-salpeter-four-point-pole-factorization}.  Here \(D(k)\) is the
exact propagator of the local field \(\phi\), including its pole residue
\(Z_\phi\).  Thus no extra \(Z_\phi\) factor is hidden in
\(\Psi_B\).  Changing to LSZ-normalized constituent fields sends
\[
  \widetilde\Phi_B\mapsto Z_\phi^{-1}\widetilde\Phi_B,
  \qquad
  D\mapsto Z_\phi^{-1}D,
  \qquad
  \Psi_B\mapsto Z_\phi\,\Psi_B
\]
for two identical scalar constituents, so the convention can be translated
without changing the pole condition.
In unamputated form the equation reads
\[
  \widetilde\Phi_B(p;P)
  =
  \mathcal G_P(p)
  \int\frac{\dd^D q}{(2\pi)^D}\,
  \mathcal K_P(p,q)\widetilde\Phi_B(q;P),
  \qquad
  P^2=-M_B^2 .
\]
This operator formulation is the pole data used later in analytic and
dispersive arguments.  When a stable composite state appears in a crossed
channel, crossing transports the pole \(P^2=-M_B^2\) to the corresponding
Mandelstam location, while
Theorem~\ref{thm:bethe-salpeter-pole-factorization} supplies the residue as a
product of Bethe--Salpeter amplitudes.  In a two-constituent channel the mass
of the pole is equivalently the value of \(P^2\) for which the integral operator
\(\widehat V_P\) has eigenvalue one.  Thus the bound-state pole terms added in
fixed-\(t\) dispersion formulae are not independent terms in a meromorphic ansatz:
their positions and residues are the spectral and Bethe--Salpeter data of the
stable one-particle sector.

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.82,
  leg/.style={line width=0.82pt},
  kern/.style={draw,fill=blue!8,rounded corners=2pt,minimum width=0.7cm,minimum height=1.25cm,inner sep=0pt},
  amp/.style={draw,fill=green!10,ellipse,minimum width=1.18cm,minimum height=1.0cm,inner sep=0pt},
  cut/.style={green!55!black,dashed,line width=0.9pt}]
  \begin{scope}[xshift=-5.75cm,yshift=1.05cm]
    \node[kern] (k) at (0,0) {\(\mathcal K\)};
    \draw[leg] (-1.1,0.42)--(k.west|-0,0.42);
    \draw[leg] (-1.1,-0.42)--(k.west|-0,-0.42);
    \draw[leg] (k.east|-0,0.42)--(1.1,0.42);
    \draw[leg] (k.east|-0,-0.42)--(1.1,-0.42);
    \node[align=center,font=\small] at (0,-1.0)
      {allowed kernel:\\no two-particle cut};
  \end{scope}
  \begin{scope}[xshift=-1.85cm,yshift=1.05cm]
    \node[kern] (a) at (-0.58,0) {\(\mathcal K\)};
    \node[kern] (b) at (0.92,0) {\(\mathcal K\)};
    \draw[leg] (-1.65,0.42)--(a.west|-0,0.42);
    \draw[leg] (-1.65,-0.42)--(a.west|-0,-0.42);
    \draw[leg] (a.east|-0,0.42)--(b.west|-0,0.42);
    \draw[leg] (a.east|-0,-0.42)--(b.west|-0,-0.42);
    \draw[leg] (b.east|-0,0.42)--(1.98,0.42);
    \draw[leg] (b.east|-0,-0.42)--(1.98,-0.42);
    \draw[cut] (0.18,-0.72)--(0.18,0.72);
    \node[orange!75!black,align=center,font=\small] at (0.18,-1.05)
      {excluded from \(\mathcal K\):\\two-particle cut};
    \node[orange!75!black] at (2.22,-0.72) {\(\times\)};
  \end{scope}
  \begin{scope}[xshift=3.6cm,yshift=1.05cm]
    \node[kern] (v) at (-1.38,0) {\(\mathcal K\)};
    \node[draw,fill=gray!12,rounded corners=2pt,minimum width=0.85cm,minimum height=1.25cm,inner sep=0pt] (t) at (0.3,0) {\(\mathcal T\)};
    \draw[leg] (-2.35,0.42)--(v.west|-0,0.42);
    \draw[leg] (-2.35,-0.42)--(v.west|-0,-0.42);
    \draw[leg] (v.east|-0,0.42)--(t.west|-0,0.42);
    \draw[leg] (v.east|-0,-0.42)--(t.west|-0,-0.42);
    \draw[leg] (t.east|-0,0.42)--(1.32,0.42);
    \draw[leg] (t.east|-0,-0.42)--(1.32,-0.42);
    \node[font=\small,blue!70!black] at (-0.54,0.77) {\(\mathcal G_P\)};
    \node[align=center,font=\small] at (-0.55,-1.05)
      {\(\mathcal T_P=\mathcal K_P+\mathcal K_P\mathcal G_P\mathcal T_P\)};
  \end{scope}
  \begin{scope}[xshift=-1.0cm,yshift=-1.8cm]
    \node[amp] (psi) at (-1.4,0) {\(\Psi_B\)};
    \node at (-0.38,0) {\(=\)};
    \node[kern] (k2) at (0.62,0) {\(\mathcal K\)};
    \draw[leg] (0.98,0.42)--(1.82,0.42);
    \draw[leg] (0.98,-0.42)--(1.82,-0.42);
    \node[amp] (psi2) at (2.62,0) {\(\Psi_B\)};
    \draw[leg] (1.82,0.42)--(psi2.west|-0,0.25);
    \draw[leg] (1.82,-0.42)--(psi2.west|-0,-0.25);
    \node[font=\small,blue!70!black] at (1.5,0.78) {\(\mathcal G_P\)};
    \node[align=left,font=\scriptsize,anchor=west] at (3.45,0)
      {homogeneous equation:\\\((\mathbf 1-\widehat V_P)\Psi_B=0\)};
  \end{scope}
\end{tikzpicture}
\end{center}

\section{Ladder Limit}

The kernel \(\mathcal K_P\) is exact when it contains all two-particle
irreducible contributions in the chosen channel. A ladder approximation keeps
only a selected low-order kernel and iterates it through \(\mathcal G_P\). For
example, a single exchanged scalar particle produces the model kernel
\[
  \mathcal K_P(p,q)
  \approx
  \frac{g_{\mathrm{ex}}^2}
       {(p-q)^2+\mu^2-\ii0},
\]
with \(\mu\) the exchanged mass and with numerator factors determined by the
spins and couplings of the fields.

The ladder sum has a controlled meaning only after the same operator topology
has been fixed.  For a selected kernel \(\mathcal K_P^{(0)}\), set
\[
  \widehat V_P^{(0)}=\mathcal K_P^{(0)}\mathcal G_P .
\]
If \(\|\widehat V_P^{(0)}\|_{\mathcal B(\mathcal X)}<1\) in a region of
complex \(P^2\), the Neumann series
\[
  \mathcal T_P^{\rm lad}
  =
  \sum_{n=0}^{\infty}(\widehat V_P^{(0)})^n\mathcal K_P^{(0)}
  =
  (\mathbf 1-\widehat V_P^{(0)})^{-1}\mathcal K_P^{(0)}
\]
converges in operator norm there.  At a bound-state pole the relevant
eigenvalue reaches \(1\), so the Neumann series itself is no longer a
convergent expansion at the pole; the pole is obtained by the meromorphic
Fredholm continuation of the resolvent.  Thus a ladder Bethe--Salpeter
eigenvalue is a theorem inside a specified compact-operator approximation, and
becomes a statement about the exact QFT only when the omitted two-particle
irreducible kernels are controlled in the same topology.

In a weakly bound massive system, let
\[
  P^0=2m+E_{\mathrm{nr}},
  \qquad |E_{\mathrm{nr}}|\ll m,
  \qquad |\vec p\,|\ll m .
\]
If the kernel is instantaneous to leading order, the relative-energy integral
through the two constituent propagators reduces the Bethe--Salpeter equation
to a Schrödinger eigenvalue problem for the equal-time wavefunction. This
nonrelativistic limit is a derived approximation to the four-point pole
condition. The relativistic equation retains the full dependence on relative
time, energy flow, and field-theoretic kernel structure.
