\chapter{Renormalizability and Local Counterterms}
\label{ch:volume-ii-renormalizability-local-counterterms}
\ConventionsEuclidean


\section{Regulated Euclidean Actions}

Let \(\phi\) be a real scalar field on \(D\)-dimensional Euclidean space.
Fix a regulator, denoted collectively by \(\Lambda\).  At finite regulator the
field is an element of a regulated field space \(\mathcal F_{\Lambda}\): for
example a finite lattice field, a finite collection of Fourier modes, or a
field paired only with test functions after a smoothing kernel has been
inserted.  The regulated functional integral has the form
\[
  Z_{\Lambda}
  =
  \int D_\Lambda^{\rm ref}\phi\,
  \exp\!\left[
    -\int \dd^Dx\,\mathcal L_{E,\Lambda}
  \right],
\]
where \(D_\Lambda^{\rm ref}\phi\) is the finite-regulator reference density of
Definition~\ref{def:regulated-scalar-integration-conventions}.  In the
alternative Gaussian-reference convention, the quadratic part of
\(\mathcal L_{E,\Lambda}\) is absorbed into \(\dd\mu_{C_\Lambda}\), and the
remaining interaction is integrated with density
\(\exp\{-L_\Lambda[\phi]\}\dd\mu_{C_\Lambda}(\phi)\).  The Euclidean
Lagrangian appearing in the path integral is written as
\[
  \mathcal L_{E,\Lambda}
  =
  \frac12 \partial_{\mu}\phi\,\partial_{\mu}\phi
  +
  \frac12 m^2\phi^2
  +
  \sum_I g_I\,\mathcal O_I(\phi).
\]
Each \(\mathcal O_I\) is a local polynomial, or more generally a local
function, of \(\phi(x)\) and finitely many derivatives
\(\partial_{\mu}\phi(x),\partial_{\mu}\partial_{\nu}\phi(x),\ldots\).
The index \(I\) includes all discrete labels required to specify the monomial:
the number of fields, the derivative multi-indices, and the internal symmetry
contraction when such indices are present.  Operators that differ by a total
derivative or by a chosen integration-by-parts convention are represented once.
The coefficient \(g_I\) is a regulated local coordinate.  The collection of
regulated coordinates includes the parameters chosen as finite inputs and the
regulator-dependent coordinates later called counterterms.

Two regulator languages will be used repeatedly.  A momentum cutoff restricts
Fourier modes to
\[
  p^2<\Lambda^2
\]
in a Euclidean invariant notation, or to a corresponding finite Brillouin zone
on a lattice.  Dimensional regularization assigns analytic meaning to
momentum integrals in
\[
  D=d-\varepsilon
\]
and removes the regulator by taking \(\varepsilon\to0\) after a subtraction
prescription has been specified.  These two languages belong to different
rows of Table~\ref{tab:regulator-integration-status-catalog}.  The cutoff
language denotes a finite-regulator field integral only after the regulated
field space and density have been supplied.  Dimensional regularization is
not a measure on \(\mathcal F_\Lambda\); it is a meromorphic assignment to the
formal perturbative graph distributions and their tensor algebra.  In both
languages the regulated coordinates are allowed to depend on the regulator or
analytic continuation parameter.

\begin{definition}[Finite-parameter renormalized family]
\label{def:finite-parameter-renormalized-family}
Choose an open set \(U\subset\R^N\) with coordinates
\[
  \lambda=(\lambda_1,\ldots,\lambda_N),
\]
and choose, for each regulator value, a map from \(U\) to the regulated
coordinate space,
\[
  b_{\Lambda}\colon U\longrightarrow \{g_I\}_{I\in\mathcal I},
  \qquad
  \lambda\longmapsto g_I(\lambda,\Lambda),
 \]
or, in dimensional regularization,
\[
  b_{\varepsilon}\colon U\longrightarrow \{g_I\}_{I\in\mathcal I},
  \qquad
  \lambda\longmapsto g_I(\lambda,\varepsilon).
\]
The coordinates \(\lambda_a\) are renormalized coordinates if they are fixed by
finite conditions on regulated correlation functions or 1PI kernels and if,
after inserting the corresponding map \(b_\Lambda\) or \(b_\varepsilon\), the
chosen class of renormalized observables has a regulator-independent limit.
\end{definition}

\begin{definition}[Counterterm coordinates and insertion regularization]
\label{def:counterterm-and-insertion-regularization}
For a fixed regulator and a fixed renormalized coordinate chart
\(\lambda\mapsto b_\Lambda(\lambda)\), there is one regulated action in the
path integral,
\[
  S_\Lambda[\phi;\lambda]
  =
  \int\dd^Dx\,\mathcal L_{E,\Lambda}
       \bigl(\phi;\{g_I(\lambda,\Lambda)\}\bigr).
\]
A \emph{counterterm} is not a second Lagrangian added to this action from
outside.  It is the regulator-dependent part of the same local coordinate map
\(b_\Lambda\), described relative to a chosen finite reference chart
\(\lambda\).  Thus the split into finite reference terms and counterterm
terms is a bookkeeping split of \(S_\Lambda[\phi;\lambda]\).

An \emph{insertion regularization} is separate data.  Before operator mixing
or continuum matching is discussed, an inserted operator is represented by a
regulated functional
\[
  \mathcal O_{a,\Lambda}^{\rm ins}(x;\phi)
\]
or equivalently by a derivative with respect to a regulated local source.  A
lattice representative, a smeared functional, a point-split expression, a
heat-kernel expression, a dimensional-regularization graph rule, and a
source-derivative prescription are different insertion regularizations.  A
renormalized operator \([O_A]_\lambda\) is the finite source coordinate or
finite linear combination obtained after the regulator-dependent insertion
map and the action-coordinate map have been fixed and the chosen continuum
limit exists.  Regularization supplies finite or meromorphic objects before
the limit; renormalization is the coordinate choice that produces finite
matched observables and operator insertions.
\end{definition}

Thus a perturbatively renormalizable family consists of a finite-dimensional
family of coordinate maps into regulated Lagrangians.  Masses, residues, and
specified 1PI vertex values at Euclidean subtraction momenta are common choices
of the coordinates \(\lambda_a\), but the definition only requires the chosen
conditions to be finite and locally independent.  Local counterterms are the
regulator-dependent local coordinates in \(b_{\Lambda}\) or
\(b_{\varepsilon}\) that make these finite conditions compatible with the
existence of the continuum limit.  The phrase ``existence of the continuum
limit'' denotes a theorem or hypothesis about a specified regulator family;
selected rigorous scalar and gauge examples, together with scalar triviality
regimes, are catalogued in
Table~\ref{tab:constructive-qft-status-catalog}.

\section{Renormalized Fields and Effective Couplings}

Let \(\Gamma_{E,\Lambda}\) be the regulated Euclidean 1PI effective action,
defined by the Euclidean Legendre construction of the preceding chapter.  A
field normalization is a regulator-dependent nonzero scalar \(Z_R(\lambda,
\Lambda)\), or a matrix in a theory with several fields.  For a single scalar
we write
\[
  \phi_R(x)=Z_R^{-1/2}\phi(x).
\]
The normalization condition can be imposed, for instance, by requiring the
coefficient of \(p^2\) in the renormalized two-point 1PI kernel at a chosen
Euclidean subtraction point to equal a prescribed finite value.  The precise
condition is part of the coordinate choice \(\lambda\).

The Green-function finiteness criterion is distributional.  For every \(n\)
in the chosen class and every test function
\(f\in C_c^\infty((\R^D)^n)\), require
\[
  \lim_{\Lambda\to\infty}
  \int\prod_{a=1}^n\dd^Dx_a\,
  f(x_1,\ldots,x_n)
  \vev{\phi_R(x_1)\cdots\phi_R(x_n)}_{E,\Lambda;\lambda}
\]
to exist.  The same statement in dimensional regularization is the existence
of the \(\varepsilon\to0\) limit after the prescribed subtractions.  This is a
criterion for the correlation distributions themselves; it is stronger than
the statement that a finite number of Taylor coefficients is finite.

On a Euclidean momentum domain where the two-point kernel is invertible and no
infrared singularity is being tested, the connected Green functions are
reconstructed from the exact 1PI kernels by trees.  Therefore, order by order
in perturbation theory, finiteness of the renormalized 1PI kernels
\[
  \Gamma^{(n)}_{E,R,\Lambda}(p_1,\ldots,p_n;\lambda)
  =
  Z_R^{n/2}
  \Gamma^{(n)}_{E,\Lambda}(p_1,\ldots,p_n;b_\Lambda(\lambda))
\]
as distributions on such momentum domains is equivalent to finiteness of the
connected functions on the corresponding domains.  This equivalence uses the
Legendre relation and the inverse two-point kernel.

For local counterterm classification one extracts only the local Taylor
coefficients of the renormalized 1PI action.  Define
\[
  \widetilde\Gamma_{E,\Lambda}[\phi_R]
  =
  \Gamma_{E,\Lambda}[Z_R^{1/2}\phi_R].
\]
Near a subtraction point at which the relevant kernels are analytic, write the
local part as
\[
  \widetilde\Gamma_{E,\Lambda}^{\rm loc}[\phi_R]
  =
  S_{\rm kin}[\phi_R]
  +
  \int \dd^Dx\,
  \sum_I g^{\rm eff}_{I,\Lambda}(\lambda)\,
  \mathcal O_I(\phi_R).
\]
Here \(S_{\rm kin}\) is the chosen finite quadratic reference action and
\(\widetilde\Gamma^{\rm loc}\) denotes the Taylor-polynomial part in external
momenta.  The local 1PI finiteness condition is
\[
  \lim_{\Lambda\to\infty}
  g^{\rm eff}_{I,\Lambda}(\lambda)
  \qquad
  \hbox{finite}
\]
for every local operator coefficient included in the finite-list
renormalized description.  Nonlocal terms in the exact 1PI action, such as
threshold logarithms, belong to the finite continuum correlation functions.

The same structure appears directly in the regulated action.  Suppose the
cutoff-dependent quadratic coordinate is represented by
\[
  S_{\rm quad}[\phi]
  =
  \frac12\int \dd^Dx\,
  \phi(-\partial^2+m_0^2)\phi .
\]
After \(\phi=Z_R^{1/2}\phi_R\), choose a finite mass parameter \(m_R\) and
write
\[
\begin{aligned}
  S_{\rm quad}[Z_R^{1/2}\phi_R]
  &=
  \frac12\int \dd^Dx\,
  \phi_R(-\partial^2+m_R^2)\phi_R     \\
  &\quad
  +
  \frac{Z_R-1}{2}
  \int \dd^Dx\,
  \phi_R(-\partial^2+m_R^2)\phi_R
  +
  \frac12\delta m^2
  \int \dd^Dx\,\phi_R^2 .
\end{aligned}
\]
The first line is the finite quadratic reference action.  The second line is
local and is treated perturbatively as the wavefunction and mass counterterm
part of the interaction.

Thus perturbation theory is organized by writing
\[
  \widetilde S_{\Lambda}[\phi_R]
  =
  S_{\rm ref}[\phi_R]
  +
  S_{\rm int}[\phi_R;\lambda]
  +
  S_{\rm ct}[\phi_R;\lambda,\Lambda],
\]
where \(S_{\rm ct}\) is local and regulator dependent.  The finite parts of
the counterterms are part of the renormalization prescription: changing them
changes the coordinate system \(\lambda\) on the same perturbative family.
The displayed split is a split of the single regulated action
\(\widetilde S_\Lambda[\phi_R]\), not a division into a physical action and an
external local functional.

\section{Locality as Extension of Distributions}

The locality of counterterms has an intrinsic position-space formulation.  In
position space, a Feynman graph with vertices
\[
  x_1,\ldots,x_r
\]
defines a distribution on the configuration space where no ultraviolet
collision has occurred.  Ultraviolet renormalization extends this distribution
across the collision diagonals, such as
\[
  x_{i_1}=\cdots=x_{i_k}.
\]
The possible ambiguities in the extension are local distributions supported on
those diagonals.

\begin{definition}[Scaling degree]
\label{def:scaling-degree}
Let \(T\in\mathcal D'(\mathbb R^n\setminus\{0\})\).  For \(\lambda>0\), define
the scaled distribution \(T_\lambda\) by
\[
  \langle T_\lambda,f\rangle
  =
  \lambda^{-n}
  \left\langle T, f(\lambda^{-1}\,\cdot)\right\rangle ,
  \qquad f\in C_c^\infty(\mathbb R^n\setminus\{0\}).
\]
The scaling degree of \(T\) at the origin is
\[
  \operatorname{sd}_0(T)
  =
  \inf\left\{\delta\in\mathbb R:
  \lambda^\delta T_\lambda\to0
  \ \hbox{in}\ \mathcal D'(\mathbb R^n\setminus\{0\})
  \ \hbox{as}\ \lambda\to0^+\right\}.
\]
\end{definition}

The extension theorem is the distribution-theoretic core of perturbative
locality.

\begin{theorem}[Extension at a diagonal]
\label{thm:extension-at-a-diagonal}
Let \(T\in\mathcal D'(\mathbb R^n\setminus\{0\})\) have finite scaling degree
\(s=\operatorname{sd}_0(T)\).  There exists an extension
\(\overline T\in\mathcal D'(\mathbb R^n)\) whose restriction to
\(\mathbb R^n\setminus\{0\}\) is \(T\) and whose scaling degree is still
\(s\).  If \(s<n\), this extension is unique.  If \(s\ge n\), any two
extensions of scaling degree \(s\) differ by
\[
  \sum_{|\alpha|\le \lfloor s\rfloor-n}
  c_\alpha\,\partial^\alpha\delta^{(n)}(x),
\]
with constants \(c_\alpha\).
\end{theorem}

\begin{proof}
We give the proof in the one-diagonal model; the general diagonal is reduced
to this model by separating center-of-mass and relative coordinates.  Let
\(m=\lfloor s\rfloor-n\).  If \(s<n\), choose a smooth cutoff \(\chi\) that is
equal to \(1\) near \(0\), and set \(\chi_\epsilon(x)=\chi(x/\epsilon)\).
For \(f\in C_c^\infty(\mathbb R^n)\), define
\[
  \langle \overline T,f\rangle
  =
  \lim_{\epsilon\to0^+}
  \langle T,(1-\chi_\epsilon)f\rangle .
\]
The scaling-degree hypothesis with \(s<n\) implies that the contribution from
the annulus \(\epsilon<|x|<2\epsilon\) is \(O(\epsilon^{n-s-\eta})\) for every
\(\eta>0\) small enough; hence the limit exists and is independent of the
cutoff.  This defines an extension with the same scaling degree.  If two such
extensions existed, their difference would be supported at \(0\).  A
distribution supported at \(0\) is a finite sum
\(\sum c_\alpha\partial^\alpha\delta\), and
\(\operatorname{sd}_0(\partial^\alpha\delta)=n+|\alpha|\ge n\).  Since
\(s<n\), no nonzero supported-at-zero difference is allowed.  This proves
existence and uniqueness in the convergent case.

Assume now \(s\ge n\).  Choose test functions \(w_\alpha\in
C_c^\infty(\mathbb R^n)\), indexed by \(|\alpha|\le m\), such that
\[
  \partial^\beta w_\alpha(0)=\delta_{\alpha\beta}
  \qquad (|\alpha|,|\beta|\le m).
\]
Define the Taylor-subtraction projector
\[
  Wf
  =
  f-\sum_{|\alpha|\le m}(\partial^\alpha f)(0)\,w_\alpha .
\]
Then all derivatives of \(Wf\) at the origin through order \(m\) vanish.  This
vanishing improves the small-scale behavior exactly enough that the same
cutoff limit
\[
  \langle \overline T_0,f\rangle
  =
  \lim_{\epsilon\to0^+}
  \langle T,(1-\chi_\epsilon)Wf\rangle
\]
exists and has scaling degree \(s\).  The functional \(\overline T_0\) agrees
with \(T\) on test functions supported away from \(0\), because then the
subtraction terms may be chosen away from the support and the cutoff is
eventually zero on the support.

The choice of \(w_\alpha\) and the values assigned to the Taylor coefficients
at the origin are not canonical.  Adding
\[
  \sum_{|\alpha|\le m} c_\alpha\,\partial^\alpha\delta^{(n)}
\]
keeps the restriction to \(\mathbb R^n\setminus\{0\}\) and has scaling degree
at most \(n+m\le s\).  Conversely, if \(\overline T_1\) and
\(\overline T_2\) are two extensions with scaling degree \(s\), their
difference is supported at \(0\), so it is a finite sum of derivatives of
\(\delta\).  The scaling-degree bound excludes all terms with
\(n+|\alpha|>s\), hence \(|\alpha|\le\lfloor s\rfloor-n=m\).  This is exactly
the displayed ambiguity.
\end{proof}

\begin{remark}[Ultraviolet content of the scaling-degree theorem]
\label{rem:uv-content-scaling-degree-theorem}
This theorem is local at the diagonal.  Its hypotheses already assume that a
distribution \(T\) is defined on the punctured configuration space
\(\mathbb R^n\setminus\{0\}\), and the conclusion controls only the missing
ultraviolet point \(0\).  It does not assert that a massless graph or
Schwinger function exists as a global distribution before infrared
renormalization, nor does it control large-distance behavior, zero-momentum
limits, or exceptional momentum configurations.

For example, a massless Euclidean propagator may have finite scaling degree at
the origin as a distribution on compactly supported test functions, while an
integrated graph containing it may fail to have an infrared-finite meaning
without an additional prescription.  If an amplitude is already infrared
divergent before the collision point is restored, then the assumption
\(T\in\mathcal D'(\mathbb R^n\setminus\{0\})\) has failed.  In massless
critical applications the extension theorem is therefore used only after an
infrared-safe setup has been specified: nonexceptional external momenta, a
finite-volume or small-mass limit, a Wilsonian cutoff construction, BPHZL
subtractions, or an equivalent prescription.  With that extra datum in place,
the scaling degree classifies the local ultraviolet ambiguity, not the
infrared existence problem.
\end{remark}

For a graph subconfiguration collapsing to a diagonal, translation invariance
removes the center-of-mass coordinate, so \(n=D(k-1)\) for \(k\) colliding
vertices.  The integer
\[
  \omega=\lfloor \operatorname{sd}_0(T)\rfloor-n
\]
is the degree of local ambiguity.  A derivative
\(\partial^\alpha\delta^{(n)}(x)\) supported on the diagonal is, after
Fourier transform, a polynomial of degree \(|\alpha|\) in the momenta entering
the collapsed subgraph.  Thus the extension theorem gives exactly the local
Taylor polynomial counterterms used in momentum-space renormalization.

Multiple collision diagonals meet along smaller diagonals.  Choosing
extensions compatibly over this stratified configuration space is the
position-space form of the subdivergence problem.  The forest formulas in the
next chapter organize the same compatibility in momentum space: each divergent
subgraph corresponds to one collision diagonal, and each counterterm is the
allowed derivative-of-delta ambiguity on that diagonal.

\section{First Counterterm Census in Four Dimensions}

The local nature of the required terms is visible in low-order scalar
examples.  Vacuum-energy terms are omitted from the census because they affect
only the zero-point functional.  Linear tadpole terms are fixed by the chosen
expansion point; in a theory without a symmetry forbidding them, their
counterterms are included among the local coordinates.  In the following
diagrams, a shaded disk labelled \(1{\rm PI}\) denotes the full
one-particle-irreducible vertex being tested after this vacuum and tadpole
bookkeeping has been imposed.

First take a \(D=4\) scalar theory with a cubic vertex
\(\int g_3\phi^3\), expanded about a point at which the one-point function has
been fixed by a tadpole condition.  The one-loop two-point graph contains the
integral
\[
  g_3^2
  \int^{\Lambda}
  \frac{\dd^4q}{(2\pi)^4}\,
  \frac{1}{(q^2+m^2)^2}
  \sim
  g_3^2\log\Lambda .
\]
It contributes to the local operator \(\phi^2\), and is canceled by a mass
counterterm.  The corresponding one-loop triangle and box integrals are
ultraviolet finite in four dimensions.

For a \(D=4\) scalar theory with a quartic vertex \(\int g_4\phi^4\), the
one-loop two-point 1PI graph is a tadpole.  Up to the normalization chosen for
the quartic vertex, it has the form
\[
  \Sigma^{(1)}(p)
  =
  C_4 g_4
  \int^\Lambda
  \frac{\dd^4q}{(2\pi)^4}\,
  \frac{1}{q^2+m^2},
\]
where \(C_4\) is a numerical combinatorial constant.  This expression is
independent of the external momentum \(p\).  It therefore contributes to the
local operator \(\phi^2\), but not to
\(\partial_\mu\phi\,\partial_\mu\phi\); equivalently the one-loop coefficient
of the wavefunction counterterm is
\[
  \delta Z^{(1)}_{\phi^4}=0 .
\]
The first logarithmically divergent momentum-dependent two-point contribution
in \(\phi^4_4\) is the two-loop sunset graph, whose Taylor coefficient at order
\(p^2\) is local and is canceled by \(\delta Z\).  The vanishing of
\(\delta Z^{(1)}_{\phi^4}\) is only a one-loop census result; it is not a
criterion for excluding the kinetic counterterm from the local chart.  The
four-point 1PI function already has a logarithmic divergence at one loop.  The
candidate \(D=4\), \(\phi^4\) local chart therefore contains
\[
  \delta m^2\phi^2,
  \qquad
  \delta Z\,\partial_{\mu}\phi\,\partial_{\mu}\phi,
  \qquad
  \delta g_4\phi^4 .
\]
The power-counting argument in the next section then proves that this candidate
list is closed: higher \(n\)-point 1PI functions with \(n\ge6\) have negative
superficial degree of divergence.

For \(D=4\) scalar theory with a sextic vertex \(\int g_6\phi^6\), the first
tests already generate both a cutoff-dependent correction to the sextic
coupling and higher local terms.  The six-point 1PI kernel receives a
quadratically divergent local contribution of the form
\[
  \delta g_6\,\phi^6,
  \qquad
  \delta g_6=O(g_6^2\Lambda^2)
\]
up to logarithms and finite prescription-dependent terms.  The eight-point
1PI kernel has logarithmic behavior,
\[
  \delta g_8\,\phi^8,
  \qquad
  \delta g_8=O(g_6^2\log\Lambda).
\]
At the next stages, the same local-subgraph mechanism generates
\[
  \delta g_{10}\phi^{10},\qquad
  \delta g_{12}\phi^{12},\qquad \ldots .
\]
The cutoff-removal limit then requires an unbounded sequence of independent
local coefficients.

Figure~\ref{fig:volume-ii-counterterm-census} summarizes this first counterterm
census: the finite local list for renormalizable scalar interactions in four
dimensions and the proliferation generated by a sextic interaction.

\begin{figure}[H]
  \centering
  \resizebox{0.97\textwidth}{!}{%
  \begin{tikzpicture}[>=Stealth,x=1cm,y=1cm,every node/.style={font=\small}]
    \tikzset{
      line/.style={line width=0.95pt},
      blob/.style={circle,draw=black,fill=gray!18,line width=0.9pt,minimum size=0.8cm}
    }
    \begin{scope}
      \node[font=\normalsize] at (0,2.55) {\(D=4,\ \phi^3\)};
      \node[blob] (A) at (-1.4,1.15) {\(1{\rm PI}\)};
      \draw[line] (-2.55,1.15)--(A.west);
      \draw[line] (A.east)--(-0.25,1.15);
      \node at (0.1,1.15) {\(=\)};
      \draw[line] (0.55,1.15)--(2.15,1.15);
      \draw[line] (0.75,1.15)..controls (1.15,1.85) and (1.75,1.85)..(1.95,1.15);
      \node[blue!70!black,align=center] at (1.35,0.3)
        {\(\sim g_3^2\log\Lambda\)\\\(\phi^2\) counterterm};
      \coordinate (u) at (-1.55,-1.1);
      \coordinate (v) at (-0.9,-0.1);
      \coordinate (w) at (-0.25,-1.1);
      \draw[line] (u)--(v)--(w)--(u);
      \draw[line] (v)--(-0.9,0.58);
      \draw[line] (u)--(-2.15,-1.7);
      \draw[line] (w)--(0.35,-1.7);
      \node[green!50!black] at (1.15,-1.1) {triangle finite};
    \end{scope}

    \begin{scope}[shift={(5.6,0)}]
      \node[font=\normalsize] at (0,2.55) {\(D=4,\ \phi^4\)};
      \coordinate (T) at (0,1.58);
      \draw[line] (-2.15,1.58)--(2.15,1.58);
      \fill (T) circle (2.4pt);
      \draw[line] (T)..controls (-0.62,2.2) and (0.62,2.2)..(T);
      \node[blue!70!black,align=center] at (0,0.88)
        {one-loop tadpole\\\(\delta m^2\), no \(p^2\)};
      \coordinate (SL) at (-1.0,-0.08);
      \coordinate (SR) at (1.0,-0.08);
      \draw[line] (-2.15,-0.08)--(SL);
      \draw[line] (SR)--(2.15,-0.08);
      \fill (SL) circle (2.4pt);
      \fill (SR) circle (2.4pt);
      \draw[line] (SL)--(SR);
      \draw[line] (SL)..controls (-0.45,0.5) and (0.45,0.5)..(SR);
      \draw[line] (SL)..controls (-0.45,-0.66) and (0.45,-0.66)..(SR);
      \node[blue!70!black,align=center] at (0,-1.08)
        {two-loop sunset\\\(\delta Z\)};
      \coordinate (BL) at (-0.78,-1.78);
      \coordinate (BR) at (0.78,-1.78);
      \fill (BL) circle (2.4pt);
      \fill (BR) circle (2.4pt);
      \draw[line] (BL)..controls (-0.25,-1.42) and (0.25,-1.42)..(BR);
      \draw[line] (BL)..controls (-0.25,-2.14) and (0.25,-2.14)..(BR);
      \draw[line] (BL)--(-1.75,-1.28);
      \draw[line] (BL)--(-1.75,-2.28);
      \draw[line] (BR)--(1.75,-1.28);
      \draw[line] (BR)--(1.75,-2.28);
      \node[green!50!black] at (0,-2.62) {one-loop \(\phi^4\) logarithm};
    \end{scope}

    \begin{scope}[shift={(11.1,0)}]
      \node[font=\normalsize] at (0,2.55) {\(D=4,\ \phi^6\)};
      \coordinate (L) at (-0.62,0.55);
      \coordinate (R) at (0.62,0.55);
      \fill (L) circle (2.3pt);
      \fill (R) circle (2.3pt);
      \draw[line] (L)..controls (-0.25,1.15) and (0.25,1.15)..(R);
      \draw[line] (L)..controls (-0.25,-0.05) and (0.25,-0.05)..(R);
      \draw[line] (L)--(-1.55,1.45);
      \draw[line] (L)--(-1.78,0.95);
      \draw[line] (L)--(-1.78,0.08);
      \draw[line] (L)--(-1.55,-0.45);
      \draw[line] (R)--(1.55,1.45);
      \draw[line] (R)--(1.78,0.95);
      \draw[line] (R)--(1.78,0.08);
      \draw[line] (R)--(1.55,-0.45);
      \node[red!70!black,align=center] at (0,-1.35)
        {six-point local term\\then \(\phi^8,\phi^{10},\ldots\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{A first counterterm census in four dimensions, after vacuum-energy
  and tadpole bookkeeping has been fixed.  Cubic and quartic scalar
  interactions require local terms from a finite operator list.  In the
  quartic theory the one-loop tadpole fixes only the mass counterterm, while
  the field-strength counterterm first appears through the two-loop sunset.
  A sextic interaction generates higher local operators under loop
  corrections.}
  \label{fig:volume-ii-counterterm-census}
\end{figure}

\section{Engineering Dimensions and the Finite List}

The preceding examples are organized by engineering dimension.  Assign
dimension \([x]=-1\) and \([\partial_\mu]=1\).  The Euclidean action is
dimensionless.  Since the kinetic term is
\[
  \frac12\int \dd^Dx\,\partial_{\mu}\phi\,\partial_{\mu}\phi ,
\]
the field has engineering dimension
\[
  [\phi]=\frac{D-2}{2}.
\]
Let \(\mathcal O_I\) contain \(n_I\) fields and \(\ell_I\) derivatives.  Its
engineering dimension is
\[
  d_I=\ell_I+n_I\frac{D-2}{2},
\]
and the coupling multiplying it has dimension
\[
  [g_I]=D-d_I .
\]

\begin{proposition}[Power-counting finite-list criterion]
\label{prop:power-counting-finite-list-criterion}
Let \(G\) be a graph contributing to the coefficient of a local operator
\(\mathcal O_I\) in the 1PI effective action.  Denote by \(V(G)\) the multiset
of interaction vertices in the graph, and let \(d_v\) be the engineering
dimension of the local operator at \(v\).  After the external momentum
derivatives that extract \(\mathcal O_I\) have been taken, the superficial
ultraviolet scaling has the form
\[
  \left(\prod_{v\in V(G)}g_v\right)
  \Lambda^{
    (D-d_I)-\sum_{v\in V(G)}(D-d_v)
  }
  \bigl(\log\Lambda\bigr)^{r},
  \qquad
  0\le r\le L,
\]
where \(L\) is the loop number.  An overall local counterterm for
\(\mathcal O_I\) can be required only when
\[
  D-d_I-\sum_{v\in V(G)}(D-d_v)\ge0 .
  \label{eq:finite-list-power-counting-inequality}
\]
If all interaction vertices present in the regulated local chart have
\(d_v\le D\), then every superficially required overall local counterterm has
\(d_I\le D\).  For scalar polynomial interactions with finitely many
derivatives at fixed \(D\), after imposing the chosen symmetries and
integration-by-parts conventions, the space of such local operators is finite
dimensional.
\end{proposition}

\begin{proof}
The coefficient \(g_I^{\rm eff}\) of \(\mathcal O_I\) in the 1PI effective
action is extracted by differentiating an \(n_I\)-point 1PI vertex with
respect to the appropriate external momenta and then setting those momenta to
zero, or to a specified Euclidean subtraction point, provided the chosen point
is infrared regular.  Under the ultraviolet rescaling of all loop momenta by
\(\Lambda\), the coefficient of a dimension-\(d_I\) local operator must have
engineering dimension \(D-d_I\).  The product of vertices contributes the
dimension \(\sum_v(D-d_v)\).  Therefore the remaining cutoff dependence has
the power displayed in the statement; logarithms can occur only when
homogeneous powers collide or after nested integrations, and their degree is
bounded at fixed loop order.  This is a statement about the overall
ultraviolet scaling after subdivergences have either been absent or treated
recursively by the next chapter's forest formula.

If every coupling appearing in the action has \(d_v\le D\), then
\(D-d_v\ge0\).  Inequality
\eqref{eq:finite-list-power-counting-inequality} implies \(d_I\le D\).  For
scalar
polynomial interactions with finitely many derivatives at fixed \(D\), after
imposing the chosen symmetries and integration-by-parts conventions there are
only finitely many local operators with \(d_I\le D\).  This is the
power-counting finite-list criterion for perturbative renormalizability: every
local ambiguity allowed by superficial ultraviolet scaling lies in a fixed
finite-dimensional local operator space.
\end{proof}

If an interaction with \(d_J>D\) is included and its coefficient is allowed to
participate in the cutoff-removal problem on the same footing, repeated
insertions can satisfy the inequality for operators of arbitrarily large
dimension.  Such a Lagrangian is still a valid low-energy effective
description at finite cutoff.  Its cutoff-removal problem, with that
interaction iterated to all orders on the same footing as marginal or relevant
couplings, requires an unbounded sequence of independent local coordinates by
this power-counting criterion.

\section{A Renormalizable Model in Six Dimensions}

The scalar cubic interaction is marginal in \(D=6\).  Use the Euclidean
Lagrangian, written in terms of the renormalized field,
\[
\begin{aligned}
  \mathcal L_E
  &=
  \frac12\partial_{\mu}\phi\,\partial_{\mu}\phi
  +
  \frac12 m_R^2\phi^2
  +
  \delta Z
  \left(
    \frac12\partial_{\mu}\phi\,\partial_{\mu}\phi
    +
    \frac12 m_R^2\phi^2
  \right)          \\
  &\quad
  +
  \frac12\delta m^2\phi^2
  +
  \frac1{3!}(g_R+\delta g)\phi^3 .
\end{aligned}
\]
In dimensional regularization with \(D=6-\varepsilon\), the last term is
understood as
\[
  \frac{\mu^{\varepsilon/2}}{3!}(g_R+\delta g)\phi^3,
\]
so that \(g_R\) is dimensionless and \(\mu\) is the subtraction scale.  The
finite parameters are \(m_R\), \(g_R\), and the normalization convention for
\(\phi\).  The counterterm coordinates are
\[
  \delta Z,\qquad \delta m^2,\qquad \delta g .
\]
They are functions of \(m_R,g_R,\mu\) and of the regulator.

\begin{proposition}[One-loop local poles in six-dimensional cubic theory]
\label{prop:six-dimensional-phi-three-one-loop-local-poles}
In the \(D=6-\varepsilon\) scalar cubic model above, the order-\(g_R^2\)
self-energy pole is a local polynomial
\[
  \Sigma_2(k)\big|_{\rm pole}
  =
  -\frac{g_R^2}{6(4\pi)^3}\frac{k^2}{\varepsilon}
  -
  \frac{g_R^2}{(4\pi)^3}\frac{m_R^2}{\varepsilon}
  +O(\varepsilon^0).
\]
It is canceled by \(\delta Z\) and \(\delta m^2\).  The order-\(g_R^3\)
one-loop three-point pole is local in the cubic vertex and is canceled by
\(\delta g\), while the one-loop four-point box is ultraviolet finite by
power counting.
\end{proposition}

\begin{proof}
At order \(g_R^2\), the one-loop self-energy is
\[
  \Sigma_{2}(k)
  =
  \frac12 g_R^2
  \int
  \frac{\dd^D\ell}{(2\pi)^D}\,
  \frac{1}
       {(\ell^2+m_R^2)((k-\ell)^2+m_R^2)} .
\]
Combining denominators gives
\[
  \Sigma_{2}(k)
  =
  \frac12 g_R^2
  \int_0^1\dd x
  \int
  \frac{\dd^D\ell}{(2\pi)^D}\,
  \frac{1}
       {\left(\ell^2+x(1-x)k^2+m_R^2\right)^2}.
\]
In dimensional regularization,
\[
  \int\frac{\dd^D\ell}{(2\pi)^D}\,
  \frac1{(\ell^2+\Delta)^n}
  =
  \frac1{(4\pi)^{D/2}}
  \frac{\Gamma(n-D/2)}{\Gamma(n)}
  \Delta^{D/2-n}.
\]
For \(D=6-\varepsilon\), this gives
\[
  \Sigma_2(k)
  =
  \frac12 g_R^2
  \frac{\Gamma(-1+\varepsilon/2)}
       {(4\pi)^{3-\varepsilon/2}}
  \int_0^1\dd x\,
  \bigl(m_R^2+x(1-x)k^2\bigr)^{1-\varepsilon/2}.
\]
The pole part is therefore a local polynomial
\[
  \Sigma_2(k)\big|_{\rm pole}
  =
  A_1(\varepsilon)\,k^2
  +
  B_1(\varepsilon)\,m_R^2,
\]
where \(A_1\) and \(B_1\) have simple \(1/\varepsilon\) poles in minimal
subtraction.  In the normalization of the displayed integral,
\[
  A_1(\varepsilon)
  =
  -\frac{g_R^2}{6(4\pi)^3}\frac1{\varepsilon}
  +O(\varepsilon^0),
  \qquad
  B_1(\varepsilon)
  =
  -\frac{g_R^2}{(4\pi)^3}\frac1{\varepsilon}
  +O(\varepsilon^0),
\]
where the finite parts depend on the subtraction prescription.  The finite
renormalized self-energy has the large-momentum form
\[
  \Sigma_{2,R}(k)
  =
  (\alpha k^2+\beta m_R^2)
  \log\frac{k^2}{\mu^2}
  +c_1k^2+c_0m_R^2+O(m_R^4/k^2)
  \qquad (k^2\gg m_R^2)
\]
for finite constants \(\alpha,\beta,c_0,c_1\) fixed by the chosen finite
renormalization.  The counterterms \(\delta Z\) and \(\delta m^2\) are chosen
so that the pole part is canceled and \(\Sigma_{2,R}\) is finite at the
subtraction point.

At order \(g_R^3\), the one-loop three-point triangle has a logarithmic
ultraviolet divergence.  Its pole is independent of external momenta at the
local order relevant for the marginal vertex, so it is canceled by
\(\delta g\).  The one-loop four-point box in \(D=6\) is ultraviolet finite by
power counting.
\end{proof}

Figure~\ref{fig:volume-ii-six-dimensional-phi-three-one-loop} displays the two
one-loop kernels used in this six-dimensional \(\phi^3\) check: the two-point
kernel with mass and kinetic local parts, and the local three-point vertex
subtraction.

\begin{figure}[H]
  \centering
  \resizebox{0.95\textwidth}{!}{%
  \begin{tikzpicture}[>=Stealth,x=1cm,y=1cm,every node/.style={font=\small}]
    \tikzset{
      line/.style={line width=0.95pt},
      blob/.style={circle,draw=black,fill=gray!18,line width=0.9pt,minimum size=0.82cm}
    }
    \node[font=\normalsize] at (0,2.0) {\(D=6,\ \phi^3\)};

    \node[blob] (A) at (-5.1,0.75) {\(1{\rm PI}\)};
    \draw[line] (-6.3,0.75)--(A.west);
    \draw[line] (A.east)--(-3.9,0.75);
    \node at (-3.55,0.75) {\(=\)};
    \draw[line] (-3.05,0.75)--(-1.7,0.75);
    \draw[line] (-2.85,0.75)..controls (-2.45,1.38) and (-2.1,1.38)..(-1.9,0.75);
    \node[blue!70!black,align=center] at (-2.35,-0.05)
      {\(A_1 k^2+B_1m_R^2\)\\\(\delta Z,\delta m^2\)};
    \node at (-0.75,0.75) {\(+\)};
    \draw[line] (-0.25,0.75)--(1.05,0.75);
    \draw[red!75!black,line width=1pt] (0.3,0.65)--(0.5,0.85);
    \draw[red!75!black,line width=1pt] (0.3,0.85)--(0.5,0.65);

    \node[blob] (B) at (-4.9,-1.4) {\(1{\rm PI}\)};
    \foreach \a in {90,210,330} {
      \draw[line] ({-4.9+0.41*cos(\a)},{-1.4+0.41*sin(\a)})
        --++({0.85*cos(\a)},{0.85*sin(\a)});
    }
    \node at (-3.55,-1.4) {\(=\)};
    \coordinate (u) at (-2.45,-0.75);
    \coordinate (v) at (-3.0,-1.75);
    \coordinate (w) at (-1.9,-1.75);
    \draw[line] (u)--(v)--(w)--(u);
    \draw[line] (u)--(-2.45,-0.05);
    \draw[line] (v)--(-3.65,-2.25);
    \draw[line] (w)--(-1.25,-2.25);
    \node[green!50!black] at (-0.1,-1.35) {\(\delta g\)};
    \node at (0.65,-1.35) {\(+\)};
    \draw[line] (1.15,-1.4)--(2.15,-1.4);
    \draw[line] (1.65,-1.4)--(1.65,-0.55);
    \draw[red!75!black,line width=1pt] (1.55,-1.5)--(1.75,-1.3);
    \draw[red!75!black,line width=1pt] (1.55,-1.3)--(1.75,-1.5);
  \end{tikzpicture}
  }
  \caption{In six-dimensional \(\phi^3\) theory, the one-loop two-point
  divergence is local in \(k^2\) and \(m_R^2\), and the one-loop three-point
  divergence is local in the cubic vertex.}
  \label{fig:volume-ii-six-dimensional-phi-three-one-loop}
\end{figure}

\section{Locality at Higher Order}

The one-loop renormalized self-energy in the six-dimensional cubic theory has
large Euclidean momentum behavior
\[
  \Sigma_R(q)
  \sim
  (\alpha q^2+\beta m_R^2)\log\frac{q^2}{\mu^2},
  \qquad
  q^2\gg m_R^2,
\]
for finite constants \(\alpha,\beta\) determined by the subtraction scheme.
When this renormalized subgraph is inserted into a larger self-energy graph,
the ultraviolet region of the remaining integral is represented by
\[
  \int
  \frac{\dd^6q}{(2\pi)^6}\,
  \frac{\Sigma_R(q)}
       {(q^2+m_R^2)^2((k-q)^2+m_R^2)} .
\]
The external momentum dependence is isolated by expanding
\[
\begin{aligned}
  \frac1{(k-q)^2+m_R^2}
  &=
  \frac1{q^2+m_R^2}
  -
  \frac{k^2-2k\cdot q}{(q^2+m_R^2)^2}       \\
  &\quad
  +
  \frac{4(k\cdot q)^2}{(q^2+m_R^2)^3}
  +
  R_3(q,k),
\end{aligned}
\]
where Taylor's formula gives \(R_3(q,k)=O(|k|^3|q|^{-5})\) for \(|q|\gg |k|,m_R\).
Because
\[
  \frac{\Sigma_R(q)}{(q^2+m_R^2)^2}
  =
  O\!\left(\frac{\log q^2}{q^2}\right),
  \qquad |q|\to\infty,
\]
the remainder contributes an integrand
\[
  \dd^6q\;O\!\left(\frac{\log q^2}{|q|^7}\right),
\]
which is ultraviolet convergent.  The displayed Taylor terms are local in the
external momentum through order \(k^2\).  After angular integration, odd powers
of \(q\) vanish and
\((k\cdot q)^2\) becomes proportional to \(k^2q^2\).  Thus the divergent part
contributes only to the \(\phi^2\) and \((\partial\phi)^2\) counterterms.

Dimensional regularization expresses the same fact algebraically.  Integrals
of the form
\[
  \int\frac{\dd^Dq}{(2\pi)^D}
  \frac{\log(q^2+\Delta)}
       {(q^2+\Delta)^n}
  =
  -\frac{\partial}{\partial n}
  \left[
    \frac1{(4\pi)^{D/2}}
    \frac{\Gamma(n-D/2)}{\Gamma(n)}
    \Delta^{D/2-n}
  \right]
\]
produce double and single poles in \(\varepsilon\).  For the self-energy
insertion, the pole part has the local form
\[
  \Sigma_{\rm insertion}(k)\big|_{\rm pole}
  =
  A_{\varepsilon}k^2+B_{\varepsilon}m_R^2,
  \qquad
  A_{\varepsilon}
  =
  \frac{a_2}{\varepsilon^2}+\frac{a_1}{\varepsilon},
  \qquad
  B_{\varepsilon}
  =
  \frac{b_2}{\varepsilon^2}+\frac{b_1}{\varepsilon}.
\]
The coefficients \(a_i,b_i\) are finite constants depending on the finite
renormalization prescription.  The pole part is polynomial in the external
momentum, so it is represented by local counterterms already present in the
six-dimensional cubic Lagrangian.
The absence of a pole of the form
\[
  \frac1{\varepsilon}\log k^2
\]
is essential: such a term would be nonlocal in position space and could not
be canceled by a local Lagrangian counterterm.  The reason it is absent is
visible before integration over Feynman parameters.  For \(D=6-\varepsilon\)
and \(n=2,3\),
\[
  \frac1{(4\pi)^{D/2}}
  \frac{\Gamma(n-D/2)}{\Gamma(n)}
  \Delta^{D/2-n}
  =
  \left(\frac{c_n}{\varepsilon}+c'_n\log\Delta\right)
  \Delta^{3-n}+O(\varepsilon^0),
\]
with constants \(c_n,c'_n\).  Differentiating with respect to \(n\), as in
the logarithmic integral above, produces
\[
  \left(
    \frac{c_{n,2}}{\varepsilon^2}
    +\frac{c_{n,1}}{\varepsilon}
  \right)\Delta^{3-n}
  +O(\varepsilon^0),
\]
because the \(\varepsilon^{-1}\log\Delta\) pieces cancel between the
derivative of the gamma function and the derivative of
\(\Delta^{D/2-n}\).

Figure~\ref{fig:volume-ii-self-energy-insertion-locality} isolates the local
\(k^0\) and \(k^2\) terms left by a renormalized self-energy insertion before
the remaining large-momentum integral is made convergent.

\begin{figure}[H]
  \centering
  \resizebox{0.98\textwidth}{!}{%
  \begin{tikzpicture}[>=Stealth,x=1cm,y=1cm,every node/.style={font=\small}]
    \tikzset{
      line/.style={line width=0.95pt},
      blob/.style={ellipse,draw=black,fill=gray!18,line width=0.9pt,minimum width=1.1cm,minimum height=0.65cm},
      arr/.style={-{Latex[length=2mm]},line width=0.9pt}
    }
    \begin{scope}
      \node[font=\normalsize,align=center] at (-0.3,2.3)
        {renormalized\\subgraph insertion};
      \draw[line] (-2.9,0.6)--(-1.4,0.6);
      \node[blob] (S) at (-0.8,0.6) {\(\Sigma_R(q)\)};
      \draw[line] (S.east)--(0.5,0.6);
      \draw[line] (0.5,0.6)--(1.5,1.05);
      \draw[line] (0.5,0.6)--(1.5,0.15);
      \node[blue!70!black,align=center] at (-0.55,-0.8)
        {\((\alpha q^2+\beta m_R^2)\log(q^2/\mu^2)\)};
      \draw[arr] (2.0,0.6)--(3.3,0.6);
    \end{scope}

    \begin{scope}[shift={(5.2,0)}]
      \node[font=\normalsize,align=center] at (0,2.3)
        {Taylor terms\\in \(k\)};
      \draw[line] (-1.65,1.05)--(1.65,1.05);
      \draw[line] (-1.65,0.15)--(1.65,0.15);
      \foreach \x/\lab in {-1.05/{k^0},0/{k^1},1.05/{k^2}} {
        \draw[red!70!black,line width=0.95pt] (\x,-0.05)--(\x,1.25);
        \node[red!70!black] at (\x,-0.35) {\(\lab\)};
      }
      \node[align=center] at (0,-1.1)
        {local pieces determine\\\(\delta m^2\) and \(\delta Z\)};
      \draw[arr] (2.0,0.6)--(3.35,0.6);
    \end{scope}

    \begin{scope}[shift={(10.6,0)}]
      \node[font=\normalsize] at (0,2.3) {finite remainder};
      \draw[line] (-1.4,0.8)..controls (-0.75,1.35) and (0.75,1.35)..(1.4,0.8);
      \draw[line] (-1.4,0.8)..controls (-0.75,0.25) and (0.75,0.25)..(1.4,0.8);
      \draw[line] (-2.4,0.8)--(-1.4,0.8);
      \draw[line] (1.4,0.8)--(2.4,0.8);
      \node[green!45!black,align=center] at (0,-0.55)
        {large-\(q\) falloff\\after subtraction};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{A renormalized subgraph can leave logarithmic large-momentum
  behavior.  Expanding the remaining propagator in the external momentum
  separates the local \(k^0\) and \(k^2\) terms from a convergent remainder.}
  \label{fig:volume-ii-self-energy-insertion-locality}
\end{figure}

\section{Schwinger Parameters and Subdivergent Regions}

The same mechanism appears geometrically in Schwinger parameters.  Consider
the two-loop self-energy graph in \(D=6\) cubic theory with two loop momenta
\(\ell_1,\ell_2\):
\[
  \frac{g_R^4}{2}
  \int
  \frac{\dd^6\ell_1}{(2\pi)^6}
  \frac{\dd^6\ell_2}{(2\pi)^6}
  P(\ell_1)P(k-\ell_1)P(\ell_2)P(k-\ell_2)P(\ell_1-\ell_2),
\]
where
\[
  P(\ell)=\frac1{\ell^2+m_R^2}.
\]
Using
\[
  P(\ell)=\int_0^{\infty}\dd\alpha\,
  \ee^{-\alpha(\ell^2+m_R^2)},
\]
the Gaussian loop integrations produce an integrand of the form
\[
  \frac1{(4\pi)^6}
  \int_0^{\infty}
  \prod_{i=1}^{5}\dd\alpha_i\,
  \frac{
    \ee^{-m_R^2\sum_i\alpha_i-k^2 Q(\alpha)}
  }
  {(\det A)^3},
\]
with
\[
  A=
  \begin{pmatrix}
    \alpha_1+\alpha_2+\alpha_5 & -\alpha_5\\
    -\alpha_5 & \alpha_3+\alpha_4+\alpha_5
  \end{pmatrix}.
\]
The function \(Q(\alpha)\) is nonnegative.  Ultraviolet regions are boundary
regions in which Schwinger parameters of a subgraph go to zero.
Explicitly,
\[
  Q(\alpha)
  =
  \alpha_2+\alpha_4
  -
  (\alpha_2,\alpha_4)
  A^{-1}
  \begin{pmatrix}
    \alpha_2\\
    \alpha_4
  \end{pmatrix}
  \ge0 .
\]
Writing
\[
  F(k^2;\alpha_1,\ldots,\alpha_5)
  =
  \frac{\exp[-k^2Q(\alpha)]}{(\det A)^3},
\]
one has the homogeneous scaling relation
\[
  \rho^6
  F\!\left(
    \frac{k^2}{\rho};
    \rho\alpha_1,\ldots,\rho\alpha_5
  \right)
  =
  F(k^2;\alpha_1,\ldots,\alpha_5).
\]

There are two one-loop subdivergent regions:
\[
  \alpha_1,\alpha_2,\alpha_5\to0,
  \qquad
  \alpha_3,\alpha_4,\alpha_5\to0 .
\]
They correspond to shrinking the left or right one-loop self-energy subgraph.
The counterterm insertions for the one-loop self-energy subtract the limiting
functions associated with these boundary regions.  At the level of the
Schwinger-parameter integrand this replacement is
\[
\begin{aligned}
  \widetilde F
  &=
  F
  -
  \lim_{\rho\to0}
  \rho^3
  F(k^2;\rho\alpha_1,\rho\alpha_2,\alpha_3,\alpha_4,\rho\alpha_5)
\\
  &\quad
  -
  \lim_{\rho\to0}
  \rho^3
  F(k^2;\alpha_1,\alpha_2,\rho\alpha_3,\rho\alpha_4,\rho\alpha_5).
\end{aligned}
\]
After these subtractions, the remaining overall ultraviolet region is the
simultaneous scaling
\[
  \alpha_1,\ldots,\alpha_5\to0 .
\]
Power counting then says that its dependence on the external momentum is at
most quadratic.  Subtracting the \(k^0\) and \(k^2\) Taylor terms leaves a
finite parameter integral.  The overall counterterm is therefore again a
linear combination of \(\phi^2\) and \((\partial\phi)^2\).
Equivalently, the finite remainder may be represented by replacing
\(\widetilde F\) with
\[
  \widetilde F_R(k^2;\alpha)
  =
  \widetilde F(k^2;\alpha)
  -
  \widetilde F(0;\alpha)
  -
  \left.
    \frac{\partial\widetilde F}{\partial k^2}
  \right|_{k^2=0}
  k^2 .
\]

Figure~\ref{fig:volume-ii-schwinger-subdivergences} translates the same locality
statement into Schwinger parameters: one-loop boundary faces are subtracted
first, and the overall boundary face then supplies the remaining local terms.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[>=Stealth,x=1cm,y=1cm,every node/.style={font=\small}]
    \tikzset{line/.style={line width=0.95pt}}
    \begin{scope}
      \node[font=\normalsize] at (0,2.15) {two-loop self-energy};
      \draw[line] (-3.0,0.5)--(-2.1,0.5);
      \draw[line] (-2.1,0.5)..controls (-1.35,1.35) and (-0.45,1.35)..(0.0,0.5);
      \draw[line] (-2.1,0.5)..controls (-1.35,-0.35) and (-0.45,-0.35)..(0.0,0.5);
      \draw[line] (0.0,0.5)..controls (0.45,1.35) and (1.35,1.35)..(2.1,0.5);
      \draw[line] (0.0,0.5)..controls (0.45,-0.35) and (1.35,-0.35)..(2.1,0.5);
      \draw[line] (2.1,0.5)--(3.0,0.5);
      \draw[orange!85!black,dashed,line width=1pt] (-2.3,-0.52) rectangle (0.18,1.52);
      \draw[green!55!black,dashed,line width=1pt] (-0.18,-0.52) rectangle (2.3,1.52);
      \node[orange!85!black] at (-1.18,-0.9)
        {\(\alpha_1,\alpha_2,\alpha_5\to0\)};
      \node[green!55!black] at (1.18,-0.9)
        {\(\alpha_3,\alpha_4,\alpha_5\to0\)};
    \end{scope}

    \begin{scope}[shift={(7.7,0)}]
      \node[font=\normalsize] at (0,2.15) {nested subtraction};
      \draw[line] (-2.1,0.7)..controls (-1.2,1.45) and (1.2,1.45)..(2.1,0.7);
      \draw[line] (-2.1,0.7)..controls (-1.2,-0.05) and (1.2,-0.05)..(2.1,0.7);
      \draw[line] (-3.0,0.7)--(-2.1,0.7);
      \draw[line] (2.1,0.7)--(3.0,0.7);
      \foreach \x in {-1.3,1.3} {
        \draw[red!70!black,line width=1pt] (\x,-0.25)--(\x,1.45);
      }
      \node[red!70!black,align=center] at (0,-0.75)
        {subgraphs first};
      \node[purple!80!black,align=center] at (0,-1.45)
        {overall \(k^0,k^2\) terms last};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{In Schwinger parameters, ultraviolet divergences are boundary
  regions.  Subdivergent one-loop regions are subtracted before the remaining
  overall boundary region is expanded in local external-momentum terms.}
  \label{fig:volume-ii-schwinger-subdivergences}
\end{figure}

The preceding analysis is the local core of perturbative renormalizability for
power-counting finite-list theories.  At all orders, the same nested structure
is organized by the Bogoliubov--Parasiuk--Hepp--Zimmermann theorem: subtract
local Taylor polynomials for divergent subgraphs in a compatible nested order,
and the remaining integrals are finite.  The next chapter develops this
subtraction structure as a systematic forest formula.
