\chapter{Scattering Kernels, LSZ, and Cluster Decomposition}
\label{ch:volume-ii-scattering-kernels}
\ConventionsLorentzian

\begin{definition}[Status of time-ordered correlators in the scattering chapters]
\label{def:scattering-time-ordered-correlator-status}
The time-ordered distributions used in this part have a fixed status.  In an
exact scattering statement, \(G_N^{\mathrm c}\) denotes the connected
time-ordered distribution associated with local Hilbert-space fields
\(\widehat\phi(x)\), obtained from the Wightman distributions by the specified
ordering and boundary-value prescription.  The input needed for LSZ is then
spectral: the two-point function has an isolated one-particle pole of mass
\(m\) and residue \(Z_\phi>0\), and the \(N\)-point time-ordered distribution
has the corresponding external pole boundary values after smearing.  No path
integral is part of the definition of the scattering operator in this
case; \(S\) remains \(\Omega_{\mathrm{out}}^*\Omega_{\mathrm{in}}\) from
Haag--Ruelle theory.

In perturbative examples, Feynman graphs, Feynman parameters, and loop
integrals denote coefficientwise contributions to a regulated source
functional \(Z[J]\).  Its connected functional is the coefficientwise
logarithm of \(Z[J]\), with the Lorentzian or Euclidean factor fixed by the
chosen regulator convention; the full source-functional status declaration is
given in Chapter~\ref{ch:volume-ii-generating-functionals-1pi}, and the BPHZ
status declaration is given in
Chapter~\ref{ch:volume-ii-subdivergences-forest-formulas}.  These are formal
or regulated calculations of the time-ordered distributions whose pole
residues may then be inserted into the LSZ formula.  Such graph calculations
do not define \(S\) and do not assert the existence of an unregulated
continuum path-integral measure.

If a Euclidean construction is used instead, the time-ordered Lorentzian
distributions are obtained only after the stated Osterwalder--Schrader
reconstruction or analytic-continuation hypotheses have been verified.  Thus a
symbol such as \(G_N^{\mathrm c}\), \(Z_\phi\), or a Feynman-parameter integral
always carries one of these statuses: exact Hilbert-space distribution,
Euclidean reconstructed distribution, or formal/regulator-dependent
perturbative coefficient.
\end{definition}

\begin{definition}[Mostly-plus Mandelstam variables]
\label{def:mostly-plus-mandelstam-variables}
Throughout the scattering and analyticity chapters, physical \(2\to2\)
momenta are written as incoming positive-energy momenta \(p_1,p_2\) and
outgoing positive-energy momenta \(p_3,p_4\), with
\[
  p_i^2=-m_i^2,
  \qquad
  p_1+p_2=p_3+p_4
\]
in the mostly-plus metric.  The Mandelstam variables are
\[
  s=-(p_1+p_2)^2,\qquad
  t=-(p_1-p_3)^2,\qquad
  u=-(p_1-p_4)^2 .
\]
They obey
\[
  s+t+u=m_1^2+m_2^2+m_3^2+m_4^2 .
\]
For equal masses \(m_i=m\), in the center-of-mass frame,
\[
  s=E_{\rm cm}^2,\qquad
  t=-2|\vec p_*|^2(1-\cos\theta),\qquad
  u=-2|\vec p_*|^2(1+\cos\theta),
\]
so physical \(s\)-channel scattering has \(s\ge4m^2\) and
\[
  -(s-4m^2)\le t\le0,\qquad
  -(s-4m^2)\le u\le0 .
\]
Positive real \(t\) or \(u\) in later fixed-\(t\) and crossing discussions is
therefore a crossed-channel timelike invariant, not a physical
\(s\)-channel scattering angle.
\end{definition}


\section{Cross-Volume Scattering Architecture}
\label{sec:cross-volume-scattering-architecture}

The scattering chapters use several inequivalent objects.  This section is a
routing convention for those objects before the detailed chapters branch
apart.  Its main rule is that an amplitude, an inclusive rate, a detector
distribution, and a resonance pole are different outputs of different
hypotheses.  A formula may connect two of them only after the hypotheses on
both sides have been stated.

\begin{definition}[Status labels for scattering constructions]
\label{def:scattering-status-labels}
In this part the following labels have fixed meanings.
\begin{enumerate}
  \item A \emph{theorem} is a statement proved from the stated Hilbert-space,
  locality, spectral, and domain hypotheses.
  \item A \emph{conditional theorem} is a theorem after additional analytic
  hypotheses have been supplied, for example a factorization, asymptotic
  completeness, or convergence assumption.
  \item A \emph{regulated construction} is a finite-cutoff or perturbative
  object with a specified regulator and subtraction prescription.
  \item \emph{Numerical evidence} is a finite-volume, lattice, bootstrap, or
  Monte Carlo datum that supports but does not replace the continuum
  construction.
  \item An \emph{open problem} is a required bridge whose hypotheses,
  convergence, or physical state-space construction are not supplied in this
  monograph.
\end{enumerate}
\end{definition}

\paragraph{Framework matrix.}
\begin{description}
  \item[Massive Haag--Ruelle/LSZ.]
  The primary objects are the in/out maps, the scattering operator, and the
  external residues used to extract amplitudes.  The input is a mass gap or
  isolated shell, local fields or almost-local creators, velocity separation,
  and distributional time
  ordering.  Under those hypotheses this is a theorem producing wave-packet
  \(S\)-matrix elements and invariant amplitudes.  It stops when charged
  sectors, confinement, or massless thresholds remove the isolated particle
  shell.
  \item[Charged QED and long-range sectors.]
  The primary objects are dressed hard states, the soft photon cloud, and an
  inclusive detector resolution.  The input is Gauss-law flux, the massless
  photon spectrum, an infrared cutoff or detector resolution, and a dressing
  or inclusive prescription.  This gives conditional perturbative and
  algebraic constructions.  The exact charged electron is not an ordinary LSZ
  particle of the massless theory; inclusive rates or dressed matrix elements
  replace bare charged \(S\)-matrix kernels.
  \item[Detector and event-shape observables.]
  The primary observables are event-shape distributions and energy-flow
  measures.  Light-ray detector products give one operator realization.  The
  input is an outgoing physical state space, a positive detector measure, a
  contact convention, and either IRC safety or a declared nonperturbative
  function.  The output is an observable distribution rather than an amplitude.
  Perturbative kernels enter only after matching, subtraction, and detector
  limits have been fixed.
  \item[Analytic \(S\)-matrix and resonances.]
  The primary objects are partial waves, dispersion relations, and
  the sheet structure in which poles, cuts, and other singularities can be
  characterized.  The input is a previously constructed amplitude, unitarity,
  crossing domain, spectral condition, and analyticity assumptions.  This
  gives a conditional analytic continuation of amplitudes.  A particular
  resonance pole is additional dynamical information, for example a zero of a
  continued inverse partial wave or a finite-volume/spectral analysis that
  fixes the continued singularity; it is not produced by analyticity alone.
  Resonance poles encode unstable states; they are not normalizable
  asymptotic particles.
  \item[Confining gauge theory and QCD factorization.]
  The primary objects are hadronic cross sections, PDFs, TMDs, jets, and
  energy correlators.  The input is color-singlet external states, operator
  definitions, factorization, Glauber status, power counting, and a detector
  or parton-density scheme.  The output is a conditional factorized
  observable.  Colored partonic amplitudes are short-distance coordinates,
  not physical external \(S\)-matrix elements.
  \item[Integrable two-dimensional scattering.]
  The primary objects are factorized scattering functions, form factors, and
  TBA data.  The input is elastic factorization, Yang--Baxter constraints,
  locality axioms for form factors, convergence, and completeness assumptions.
  The output is exact algebraic and thermodynamic data.  Reconstruction of
  local observables and local algebras is a separate analytic bridge.
\end{description}

\paragraph{Massive route.}
In a massive theory with a stable isolated one-particle shell, the route is:
local fields and spectrum; Haag--Ruelle wave operators
\(\Omega_{\rm in/out}\); the scattering operator \(S\); the invariant
amplitude \(\mathcal M\); and finally the cross section \(\sigma\).
The first arrow is Haag--Ruelle theory
(Chapters~\ref{ch:haag-ruelle} and
\ref{ch:volume-iv-haag-ruelle-mathematical-scattering}).  The second is the
definition of the scattering operator.  The third is LSZ residue extraction
(Chapter~\ref{ch:lsz-reduction} and this chapter).  The fourth is phase-space
integration with the flux and normalization conventions of
Chapter~\ref{ch:cross-sections-unitarity}.  Analytic continuation, partial
waves, and resonance poles may then be studied in
Chapters~\ref{ch:volume-ii-resonances-dressed-propagators} and
\ref{ch:volume-ii-07}, but those analytic objects start from the amplitude
already obtained on the physical sheet.

\paragraph{Charged QED route.}
In massless QED the charged sector does not satisfy the massive LSZ input.
Gauss-law flux and arbitrarily soft photons change the spectral support of a
charged state, so the bare electron Green-function residue is not an exact
one-particle \(S\)-matrix coordinate of the full theory.  The route is
instead: a hard charged process; soft dressing or an inclusive sum; and a
detector-resolved rate.
Chapter~\ref{ch:qed-external-states} fixes external-state conventions for
regulated perturbative calculations.  Chapter~\ref{ch:volume-ii-infrared-inclusive-qed}
then supplies the inclusive and dressed replacements.  A perturbative hard
amplitude remains useful, but only as an ingredient in the resolution
dependent observable after the soft sector has been attached.

\paragraph{Detector and QCD route.}
For QCD event shapes and energy correlators the physical object is a
color-singlet detector distribution.  The route is: a current or hadron
source; factorized operator data; jet, soft, or parton-density functions; and
the detector measure.
Chapter~\ref{ch:volume-ii-19} gives the QCD operator and factorization
layers, Chapter~\ref{ch:jets-ir-safe-observables-hadronization} develops
event shapes, track functions, and hadronization coordinates, and
Chapter~\ref{ch:cft-light-ray-energy-correlators} supplies the conformal
light-ray and detector-algebra comparison.  A partonic amplitude or splitting
kernel is a short-distance coordinate inside this route.  It becomes a
physical prediction only after the observable, subtraction scheme, Glauber
status, endpoint/contact convention, and nonperturbative inputs have been
declared.

\paragraph{Map and non-map ledger.}
The following implications are used in later chapters.
\begin{enumerate}
  \item Haag--Ruelle plus LSZ hypotheses give massive wave-packet amplitudes.
  \item Massive amplitudes plus unitarity, boundedness, and analyticity
  hypotheses give partial-wave and dispersion-relation machinery, together
  with a continuation framework in which resonance poles, cuts, and other
  singularities can be characterized if present.  The existence and location
  of a resonance pole require separate channel-specific dynamical or spectral
  evidence.
  \item A charged hard process plus soft or inclusive data gives
  infrared-safe charged rates.
  \item QCD operator definitions plus factorization data give detector and
  event-shape distributions.
\end{enumerate}
The converse arrows are absent without extra hypotheses.  A detector
distribution does not determine a unique amplitude; a resonance pole does not
produce an asymptotic particle vector; an inclusive QED rate does not restore
a bare charged LSZ state; a finite partonic or lattice check does not supply
the continuum factorization or reconstruction theorem.  This ledger is the
cross-volume convention used whenever the later chapters compare these
languages.

\section{Scattering Operator and Kernel Notation}

This chapter fixes the distributional kernel notation used in the analytic
chapters.  The nonperturbative construction of \(S\) by Haag--Ruelle wave
operators was given in Chapters~\ref{ch:haag-ruelle} and
\ref{ch:volume-iv-haag-ruelle-mathematical-scattering}; it is not being
redefined here.

Work in one massive Haag--Ruelle scattering sector satisfying the hypotheses of
Theorem~\ref{thm:algebraic-haag-ruelle-scattering}, with the scalar
point-field construction and Cook estimates reviewed in
Proposition~\ref{prop:haag-ruelle-cook-limits}.  Fix a stable scalar
species of mass \(m>0\).  Its positive mass shell is
\[
  \Sigma_m^+
  =
  \{(\omega_{\vec p},\vec p):\vec p\in\mathbb R^d\},
  \qquad
  \omega_{\vec p}=\sqrt{\vec p^{\,2}+m^2}.
\]
The corresponding one-particle Hilbert space is denoted by \(\Hilb_1\).  In
the scalar case it is represented as \(L^2(\Sigma_m^+,\dd\mu_m)\), where
\[
  \dd\mu_m(\vec p)
  =
  \frac{\dd^d\vec p}{(2\pi)^d2\omega_{\vec p}} .
\]
The symmetric Fock space over \(\Hilb_1\) is
\[
  \mathcal F_s(\Hilb_1)
  =
  \bigoplus_{n=0}^\infty \Hilb_1^{\odot n}.
\]
The Haag--Ruelle wave operators from the preceding theorem layers
are isometries
\[
  \Omega_{\mathrm{in/out}}:
  \mathcal F_s(\Hilb_1)\longrightarrow \Hilb .
\]
On a scattering sector for which the incoming and outgoing ranges coincide,
the scattering operator is
\[
  S=\Omega_{\mathrm{out}}^*\Omega_{\mathrm{in}}
  :
  \mathcal F_s(\Hilb_1)\longrightarrow\mathcal F_s(\Hilb_1).
\]
This is the object whose kernels will be studied throughout this part.

Generalized momentum states on \(\Sigma_m^+\) are normalized by
\[
  \langle\vec q\,|\vec p\rangle
  =
  (2\pi)^d2\omega_{\vec p}\,\delta^{(d)}(\vec q-\vec p).
\]
For outgoing momenta \(q_i=(\omega_{\vec q_i},\vec q_i)\) and incoming
momenta \(p_j=(\omega_{\vec p_j},\vec p_j)\), write
\[
  S(q_1,\ldots,q_m|p_1,\ldots,p_n)
  =
  {}_{\mathrm{out}}\langle
  \vec q_1,\ldots,\vec q_m
  |\vec p_1,\ldots,\vec p_n
  \rangle_{\mathrm{in}} .
\]
This notation denotes a distributional kernel. Its value on wave packets
\(f_i,g_j\in C_c^\infty(\mathbb R^d)\) is
\[
\begin{aligned}
&S[f_1,\ldots,f_m;g_1,\ldots,g_n]
\\
&=
  \int
  \prod_{i=1}^m\dd\mu_m(\vec q_i)\,f_i(\vec q_i)^*
  \prod_{j=1}^n\dd\mu_m(\vec p_j)\,g_j(\vec p_j)
  S(q_1,\ldots,q_m|p_1,\ldots,p_n).
\end{aligned}
\]

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.98]
  \draw[->,line width=0.75pt] (-4.2,-2.0)--(-4.2,2.25)
    node[above] {$x^0$};
  \draw[orange!85!black,dashed,line width=0.9pt] (-2.35,-1.18)--(2.25,-1.18);
  \draw[green!50!black,dashed,line width=0.9pt] (-2.35,1.18)--(2.25,1.18);
  \node[orange!85!black,anchor=west] at (2.4,-1.18)
    {incoming Fock basis};
  \node[green!50!black,anchor=west] at (2.4,1.18)
    {outgoing Fock basis};
  \draw[blue!75!black,line width=1.2pt,fill=blue!5] (0,0) ellipse (1.15 and 0.74);
  \node[blue!75!black] at (0,0) {\(S\)};
  \foreach \x in {-0.78,0,0.78} {
    \draw[orange!85!black,line width=1.05pt,->]
      (\x,-1.82)--({0.48*\x},-0.62);
  }
  \foreach \x in {-0.78,0,0.78} {
    \draw[green!50!black,line width=1.05pt,->]
      ({0.48*\x},0.62)--(\x,1.82);
  }
\end{tikzpicture}
\end{center}

\section{Connected Kernels and Invariant Amplitudes}

Let
\[
  P_{\mathrm{out}}=\sum_{i=1}^m q_i,
  \qquad
  P_{\mathrm{in}}=\sum_{j=1}^n p_j .
\]
Translation invariance implies that the non-identity part of every connected
scattering kernel is supported on \(P_{\mathrm{out}}=P_{\mathrm{in}}\).  The
phase convention is fixed once and for all by distinguishing the connected
\(S\)-kernel from the connected \(T\)-kernel.  With
\[
  S=\mathbf 1+\ii T,
\]
write
\[
\begin{aligned}
&S^{\mathrm c}(q_1,\ldots,q_m|p_1,\ldots,p_n)
\\
&\qquad =
  I^{\mathrm c}(q_1,\ldots,q_m|p_1,\ldots,p_n)
  +
  \ii\,T^{\mathrm c}(q_1,\ldots,q_m|p_1,\ldots,p_n),
\end{aligned}
\]
where \(I^{\mathrm c}\) is zero except for one incoming and one outgoing
particle of the same species, in which case
\[
  I^{\mathrm c}(q|p)
  =
  (2\pi)^d2\omega_{\vec p}\delta^{(d)}(\vec q-\vec p).
\]
The invariant amplitude \(\mathcal M\) is the reduced connected \(T\)-kernel:
\[
\begin{aligned}
&T^{\mathrm c}(q_1,\ldots,q_m|p_1,\ldots,p_n)
\\
&\qquad =
  (2\pi)^D\delta^{(D)}(P_{\mathrm{out}}-P_{\mathrm{in}})
  \mathcal M(q_1,\ldots,q_m|p_1,\ldots,p_n).
\end{aligned}
\]
Equivalently, the transition part of the connected \(S\)-kernel is
\[
  S^{\mathrm c}-I^{\mathrm c}
  =
  \ii(2\pi)^D\delta^{(D)}(P_{\mathrm{out}}-P_{\mathrm{in}})\mathcal M .
\]
The connected \(T\)-matrix kernel is therefore
\[
  {}_{\mathrm{out}}\langle q|T|p\rangle_{\mathrm c}
  =
  (2\pi)^D\delta^{(D)}(P_{\mathrm{out}}-P_{\mathrm{in}})
  \mathcal M(q|p).
\]

The word connected has a recursive distributional meaning. If the full
process splits into independent subprocesses, the full kernel is the product
of the kernels of those subprocesses, with separate momentum conservation in
each component. The connected kernel is the part that remains after subtracting
all such products.

\section{LSZ as Pole Extraction}

The pole-extraction theorem is Theorem~\ref{thm:lsz-wave-packet} of
Chapter~\ref{ch:lsz-reduction}; it is not restated here.  In this chapter we
use only its kernel consequence: after wave-packet smearing, a connected
time-ordered distribution whose external variables have isolated
one-particle poles of residue \(Z_\phi\) gives the connected scattering kernel
by multiplying each external leg by \(Z_\phi^{-1/2}\ii(k_a^2+m^2)\), taking
the simultaneous external pole boundary value, and then setting outgoing
variables \(k_i=q_i\) and incoming variables \(k_{m+j}=-p_j\) on
\(\Sigma_m^+\).  The resulting distribution is a matrix element of the
already-defined Haag--Ruelle operator
\(S=\Omega_{\rm out}^*\Omega_{\rm in}\); in perturbative sections the same
operation is applied coefficientwise to regulated time-ordered kernels.

\section{Cluster Decomposition}

Let \(\alpha=\{p_1,\ldots,p_n\}\) be the incoming labels and
\(\beta=\{q_1,\ldots,q_m\}\) the outgoing labels. A process partition
\(\Pi\) is a finite set of nonempty blocks \(B\), where each block contains a
subset \(\alpha_B\subset\alpha\) and a subset \(\beta_B\subset\beta\), and
the blocks partition all incoming and outgoing labels. Define
\[
  S_B^{\mathrm c}
  =
  S^{\mathrm c}(\beta_B|\alpha_B).
\]

\paragraph{Connected scattering kernels.}
For fixed incoming and outgoing labels, the full scattering kernel has a unique
expansion
\begin{equation}
  S(\beta|\alpha)
  =
  \sum_{\Pi}
  \prod_{B\in\Pi}S^{\mathrm c}(\beta_B|\alpha_B),
  \label{eq:scattering-kernel-cluster-partition}
\end{equation}
where the sum is over process partitions.  Equivalently, the connected kernel
is determined recursively by
\begin{equation}
  S^{\mathrm c}(\beta|\alpha)
  =
  S(\beta|\alpha)
  -
  \sum_{\Pi\ne\{\alpha\cup\beta\}}
  \prod_{B\in\Pi}S^{\mathrm c}(\beta_B|\alpha_B).
  \label{eq:connected-scattering-recursion}
\end{equation}
The one-particle identity kernel \(I^{\mathrm c}\) introduced above is included
as the connected kernel for a block with one incoming and one outgoing label:
\begin{equation}
  \begin{aligned}
  S^{\mathrm c}(q|p)
  &=
  I^{\mathrm c}(q|p)
  +
  \ii(2\pi)^D\delta^{(D)}(q-p)\mathcal M(q|p)
  \\
  &=
  (2\pi)^d2\omega_{\vec p}\delta^{(d)}(\vec q-\vec p)
  +
  \ii(2\pi)^D\delta^{(D)}(q-p)\mathcal M(q|p).
  \end{aligned}
  \label{eq:one-particle-connected-kernel}
\end{equation}
In a translationally invariant vacuum without external backgrounds, the
nontrivial one-particle amplitude vanishes for a stable isolated species, and
the identity term remains.

Order the finite label sets by their total number of external labels.  For one
block the statement defines \(S^{\mathrm c}\).  Suppose the connected kernels
have been defined for every proper subcollection.  Then
\eqref{eq:connected-scattering-recursion} defines the connected kernel for the
given collection by subtracting all factorizations into proper subprocesses.
Substituting this definition back into the right-hand side of
\eqref{eq:scattering-kernel-cluster-partition} gives \(S(\beta|\alpha)\);
uniqueness follows because every other solution would have a difference whose
highest-size nonzero component appears only in the one-block term and hence
must vanish.  This is the moment-cumulant recursion for scattering kernels,
with process partitions replacing ordinary set partitions.

This recursion is algebraic: it defines
connected kernels once the full kernel is known.  The physical cluster
property is the large-separation theorem below.  Its hypotheses are additional
Hilbert-space and locality hypotheses, not consequences of the recursion.

\begin{theorem}[Cluster decomposition with a joint Haag--Ruelle limit]
\label{thm:scattering-cluster-decomposition}
Assume a Haag--Ruelle scattering theory with a unique translation-invariant
vacuum, a mass gap above the vacuum, local commutativity, and the following
quantitative clustering estimate for the almost-local Haag--Ruelle
approximants used to construct the scattering states.  If \(F_T\) and \(G_T\)
are finite products of such approximants at fixed time parameter \(T\), and
if \(G_T(a)=U(a)G_TU(a)^*\), then for every \(N\) there are constants
\(C_N,M_N,R_N\), depending on the chosen operator families and wave packets
but independent of \(T\) and \(a\), such that
\begin{equation}
  \left|
  \bra{\vac}F_TG_T(a)\ket{\vac}
  -
  \bra{\vac}F_T\ket{\vac}\bra{\vac}G_T\ket{\vac}
  \right|
  \le
  C_N(1+|T|)^{M_N}(1+\rho(a,T))^{-N}.
  \label{eq:hr-quantitative-cluster-estimate}
\end{equation}
Here \(\rho(a,T)\) is the spacelike separation of the two localization tubes
after replacing the almost-local operators by local approximants; for
velocity supports separated from the light cone of the relative translation,
\(\rho(a,T)\ge c|a|-R_N(1+|T|)\) for some \(c>0\).  Let
\(\psi_A^{\mathrm{in/out}}\) and \(\psi_B^{\mathrm{in/out}}\) be two finite
asymptotic Fock vectors with compact velocity supports.  Write
\[
  \mathsf J_{A|B}:
  \mathcal F_A\otimes\mathcal F_B\longrightarrow\mathcal F_s(\Hilb_1)
\]
for the symmetrized concatenation map on the fixed cluster-label
decomposition, so that
\(\mathsf J_{A|B}(\psi_A\otimes\psi_B)=\psi_A\odot\psi_B\).  Let \(U_F(a)\)
be the free Fock translation.  The partial relative translation of the
\(B\)-cluster is the typed vector
\begin{equation}
  \psi_{A|B}(a)
  =
  \mathsf J_{A|B}\bigl(\psi_A\otimes U_F(a)\psi_B\bigr),
  \label{eq:partial-fock-cluster-translation}
\end{equation}
not a global physical translation of the already-combined scattering state.

Assume, in addition, the following joint Haag--Ruelle approximation estimate
for this family of partially translated cluster vectors.  Let
\(W_{\sigma,T}\), \(\sigma=\mathrm{in},\mathrm{out}\), denote the finite-time
Haag--Ruelle product map corresponding to the chosen wave packets.  There are
diagonal times \(T=T(|a|)\to\infty\), with \(T(|a|)=o(|a|)\), and error
functions \(E_\sigma(T,a)\) such that
\begin{equation}
  \left\|
    W_{\sigma,T(a)}\psi_{A|B}^{\sigma}(a)
    -
    \Omega_\sigma\psi_{A|B}^{\sigma}(a)
  \right\|
  \le
  E_\sigma(T(a),a)
  \longrightarrow0 ,
  \label{eq:joint-hr-cluster-approximation}
\end{equation}
the finite-time vectors are bounded on this diagonal, and
\begin{equation}
  \rho(a,T(a))\to\infty,\qquad
  C_N(1+T(a))^{M_N}(1+\rho(a,T(a)))^{-N}\to0
  \label{eq:cluster-diagonal-decay}
\end{equation}
for some \(N\) large enough for the chosen packets.  Then the full scattering
matrix element factorizes:
\[
  \left\langle
    \psi_{A|B}^{\mathrm{out}}(a),
    S\,
    \psi_{A|B}^{\mathrm{in}}(a)
  \right\rangle
  \longrightarrow
  \left\langle\psi_A^{\mathrm{out}},S\psi_A^{\mathrm{in}}\right\rangle
  \left\langle\psi_B^{\mathrm{out}},S\psi_B^{\mathrm{in}}\right\rangle .
\]
\end{theorem}

\begin{proof}
Use \(S=\Omega_{\rm out}^*\Omega_{\rm in}\) to write the exact matrix element
as
\[
  \mathcal M(a)
  =
  \left\langle
    \Omega_{\rm out}\psi_{A|B}^{\rm out}(a),
    \Omega_{\rm in}\psi_{A|B}^{\rm in}(a)
  \right\rangle_{\Hilb}.
\]
The partial translation in
\eqref{eq:partial-fock-cluster-translation} acts on only one tensor factor
before symmetrized concatenation.  Therefore the required approximation is
\eqref{eq:joint-hr-cluster-approximation}, an additional joint input beyond
the physical unitary \(U(a)\) acting on a whole state.

Along the diagonal \(T=T(a)\), define the finite-time matrix element
\[
  M_T(a)
  =
  \left\langle
    W_{{\rm out},T}\psi_{A|B}^{\rm out}(a),
    W_{{\rm in},T}\psi_{A|B}^{\rm in}(a)
  \right\rangle_{\Hilb}.
\]
The assumed joint estimate and boundedness give
\[
  |\mathcal M(a)-M_{T(a)}(a)|
  \longrightarrow0 .
\]
For this chosen \(T\), the finite-time expression has the vacuum expectation
form
\[
  M_T(a)
  =
  \bra{\vac}
    (A_T^{\mathrm{out}})^*
    (B_T^{\mathrm{out}}(a))^*
    A_{-T}^{\mathrm{in}}
    B_{-T}^{\mathrm{in}}(a)
  \ket{\vac}.
\]
Almost-locality permits replacement of each approximant by a local operator
localized in its Haag--Ruelle tube, with an error decreasing faster than any
power of the tube size.  Along the diagonal, \(T(a)=o(|a|)\) and the lower
bound on \(\rho(a,T(a))\) make the \(A\)-tube spacelike separated from the
translated \(B\)-tube for large \(|a|\).  Local commutativity then orders the
two groups, up to the almost-local tail estimate, into
\[
  M_T(a)
  =
  \bra{\vac}F_TG_T(a)\ket{\vac}
  +
  O_N((1+|a|)^{-N}),
  \qquad
  F_T=(A_T^{\mathrm{out}})^*A_{-T}^{\mathrm{in}},
  \quad
  G_T=(B_T^{\mathrm{out}})^*B_{-T}^{\mathrm{in}} .
\]
Applying \eqref{eq:hr-quantitative-cluster-estimate} along the same diagonal,
together with \eqref{eq:cluster-diagonal-decay}, gives
\[
  M_{T(a)}(a)
  -
  \bra{\vac}F_{T(a)}\ket{\vac}\bra{\vac}G_{T(a)}\ket{\vac}
  \longrightarrow0 .
\]
Since \(T(a)\to\infty\), ordinary Haag--Ruelle convergence for each
untranslated cluster gives
\[
  \bra{\vac}F_{T(a)}\ket{\vac}
  \longrightarrow
  \left\langle\psi_A^{\mathrm{out}},S\psi_A^{\mathrm{in}}\right\rangle,
  \qquad
  \bra{\vac}G_{T(a)}\ket{\vac}
  \longrightarrow
  \left\langle\psi_B^{\mathrm{out}},S\psi_B^{\mathrm{in}}\right\rangle .
\]
Combining the three estimates proves the factorization of the exact scattering
matrix element.
\end{proof}

\paragraph{Kernel consequence and low-dimensional boundary.}
The theorem is a wave-packet statement.  Its distributional consequence is
that, in a regular massive kinematic region and in a topology obtained by
testing against compactly supported mass-shell wave packets, the connected
kernel has no component which is a product of proper subprocess kernels with
separate conservation laws.  In dimensions and kinematic regions where the
total momentum-conservation surface is a clean codimension-\(D\) constraint
and no additional elastic graph degeneracy is present, the transition part of
the connected kernel may then be written as one total
\(\delta^{(D)}(P_{\rm out}-P_{\rm in})\) times a reduced distribution.  This
last sentence is a regular-kernel corollary, not the content of the
Haag--Ruelle cluster theorem itself.

The qualification is essential in \(1+1\)-dimensional elastic scattering.  In
rapidity variables, \(p^\pm=m\ee^{\pm\theta}\).  For a two-particle elastic
process, total energy-momentum conservation gives
\[
  \ee^{\theta'_1}+\ee^{\theta'_2}
  =
  \ee^{\theta_1}+\ee^{\theta_2},
  \qquad
  \ee^{-\theta'_1}+\ee^{-\theta'_2}
  =
  \ee^{-\theta_1}+\ee^{-\theta_2}.
\]
Away from coincident rapidities this says that
\(\{\ee^{\theta'_1},\ee^{\theta'_2}\}
=\{\ee^{\theta_1},\ee^{\theta_2}\}\).  Thus the total conservation
distribution itself pulls back to the union of the identity and exchange
permutation graphs in rapidity space, the support structure used in
Proposition~\ref{prop:higher-spin-factorization-kinematic}.  Cluster
subtraction removes proper product blocks; it does not erase elastic
permutation-graph support compatible with total conservation.

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.95,
  blob/.style={ellipse,draw=black,thick,minimum width=1.45cm,
    minimum height=0.95cm,fill=orange!8}]
  \node[blob] (A) at (0,0.72) {\(S^{\mathrm c}_1\)};
  \node[blob] (B) at (0,-0.72) {\(S^{\mathrm c}_2\)};
  \draw[thick,->] (-2.2,1.2)--(A.west);
  \draw[thick,->] (-2.2,0.45)--(A.west);
  \draw[thick,->] (A.east)--(2.2,1.0);
  \draw[thick,->] (-2.2,-0.45)--(B.west);
  \draw[thick,->] (B.east)--(2.2,-0.35);
  \draw[thick,->] (B.east)--(2.2,-1.2);
  \node[left] at (-2.25,1.2) {$p_1$};
  \node[left] at (-2.25,0.45) {$p_2$};
  \node[left] at (-2.25,-0.45) {$p_3$};
  \node[right] at (2.25,1.0) {$q_1$};
  \node[right] at (2.25,-0.35) {$q_2$};
  \node[right] at (2.25,-1.2) {$q_3$};
  \node at (3.15,0) {$\Longrightarrow$};
  \node[ellipse,draw=black,thick,minimum width=1.65cm,
    minimum height=1.3cm,fill=blue!6] (C) at (4.55,0) {\(S\)};
  \foreach \y in {1.05,0.35,-0.35} {
    \draw[thick,->] (3.65,\y)--(C.west);
  }
  \foreach \y in {1.05,0,-1.05} {
    \draw[thick,->] (C.east)--(5.85,\y);
  }
\end{tikzpicture}
\end{center}

The same decomposition appears in connected Green functions. Since LSZ acts
on external pole residues, it maps the connected part of a time-ordered Green
function to the connected part of the scattering kernel with the same
external labels.

\section{Preparation for Analytic Structure}

The object carried forward is the connected invariant amplitude
\[
  \mathcal M(q_1,\ldots,q_m|p_1,\ldots,p_n)
\]
as a distributional boundary value on the physical mass shells. Bound states
will appear as poles of this object in an appropriate channel. Resonances will
appear as poles reached by analytic continuation through a continuum cut.
Renormalization will be formulated as the construction of finite Green
functions and finite amplitudes while preserving the local form of the
counterterms.

For sectors with massless long-range forces, confinement, or infraparticle
spectra, the asymptotic data require modified observables or modified state
spaces. The later chapters return to those domains after the massive
scattering and renormalization structures have been developed. In particular,
one must not identify the charged electron labels used in perturbative QED
hard kernels with exact on-shell LSZ external particles of massless QED.  The
ordinary LSZ formula requires an isolated charged one-particle pole; the QED
infrared chapters instead construct resolution-inclusive probabilities and
softly dressed charged asymptotic data.
