\chapter{Resonances and Dressed Propagators}
\label{ch:volume-ii-resonances-dressed-propagators}
\ConventionsLorentzian


\section{Above-Threshold Exchange}

Continue the two-scalar model of the preceding chapter,
\[
  \mathcal L
  =
  -\frac12\partial_\mu\phi_1\partial^\mu\phi_1
  -\frac12m_1^2\phi_1^2
  -\frac12\partial_\mu\phi_2\partial^\mu\phi_2
  -\frac12m_2^2\phi_2^2
  -\frac12g\phi_1^2\phi_2 ,
\]
with \(m_1,m_2>0\) and \(g\in\R\).

\begin{definition}[Above-threshold scalar exchange datum]
\label{def:above-threshold-scalar-exchange-datum}
The above-threshold exchange datum consists of the preceding two-scalar local
Lagrangian, stable external \(\phi_1\) one-particle states, the mass
inequality
\[
  m_2>2m_1 ,
\]
and the mostly-plus Mandelstam invariant
\[
  K=k_1+k_2,\qquad s=-K^2,
\]
where \(k_1,k_2\) are the incoming \(\phi_1\) momenta, in the convention of
Definition~\ref{def:mostly-plus-mandelstam-variables}.  The \(s\)-channel tree
contribution to the connected invariant amplitude is
\[
  \mathcal M_s^{\mathrm{tree}}(s)
  =
  \frac{g^2}{m_2^2-s-\ii0}.
\]
\end{definition}

For the datum of
Definition~\ref{def:above-threshold-scalar-exchange-datum},
the value \(s=m_2^2\) lies in the two-\(\phi_1\) scattering region
\(s\ge4m_1^2\). The relevant scattering problem is therefore a problem with
stable external \(\phi_1\) states and an open two-particle continuum in the
same channel as the exchanged \(\phi_2\) line. The exact physical-axis
\(S\)-matrix is a bounded unitary boundary value; the pole suggested by the
tree graph is the lowest-order expansion of an analytic structure reorganized
by the exact two-point function of \(\phi_2\).

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.94,
  one/.style={line width=0.85pt},
  two/.style={dash pattern=on 3pt off 2pt,line width=0.95pt},
  cut/.style={red!75!black,line width=1.6pt}]
  \begin{scope}
    \draw[one,->] (-1.65,0.9)--(-0.45,0.18);
    \draw[one,->] (-1.65,-0.9)--(-0.45,-0.18);
    \draw[two] (-0.45,0)--(1.15,0);
    \draw[one,->] (1.15,0.18)--(2.35,0.9);
    \draw[one,->] (1.15,-0.18)--(2.35,-0.9);
    \node at (0.35,0.35) {\(\phi_2\)};
    \node at (0,-1.35) {\(K^2=-s\)};
  \end{scope}
  \begin{scope}[xshift=5.1cm]
    \draw[->,line width=0.75pt] (-0.35,0)--(4.4,0) node[right] {\(s\)};
    \draw[->,line width=0.75pt] (0,-1.15)--(0,1.75)
      node[above] {\(\operatorname{Im}s\)};
    \draw[cut] (1.55,0)--(4.1,0);
    \node[red!75!black,below] at (1.55,-0.12) {\(4m_1^2\)};
    \fill[orange!90!black] (2.85,0) circle (2.4pt);
    \node[orange!90!black,below] at (2.85,-0.18) {\(m_2^2\)};
    \node[red!75!black,align=center] at (3.72,1.05)
      {open\\continuum};
    \draw[red!75!black,->,line width=0.8pt] (3.72,0.68)--(3.62,0.11);
    \draw[orange!90!black,->,line width=0.8pt] (2.85,0.65)--(2.85,0.12);
  \end{scope}
\end{tikzpicture}
\end{center}

The partial-wave normalization used below is the unordered identical-boson
convention for the physical \(\phi_1\phi_1\to\phi_1\phi_1\) channel:
\begin{equation}
  \mathcal M(s,z)
  =
  32\pi
  \sum_{\substack{\ell=0\\ \ell\ {\rm even}}}^{\infty}
  (2\ell+1)a_\ell(s)P_\ell(z),
  \qquad z=\cos\theta,
  \label{eq:resonance-partial-wave-expansion}
\end{equation}
This is the \(32\pi\), even-\(\ell\) convention of
Equation~\eqref{eq:identical-partial-wave-conversion}.  The ordered
\(16\pi\) convention may still be used as an intermediate labelled-leg
bookkeeping device, but the elastic \(S_\ell\) below acts on the symmetric
two-\(\phi_1\) Hilbert-space channel, so the physical partial-wave coordinate
is \(a_\ell\).
In an elastic one-channel interval,
\begin{equation}
  S_\ell(s)=1+2\ii\,\rho_1(s)a_\ell(s),
  \qquad
  \rho_1(s)=\sqrt{1-\frac{4m_1^2}{s}} .
  \label{eq:resonance-elastic-s-matrix-normalization}
\end{equation}
The branch of \(\rho_1\) is chosen by \(\rho_1(s)>0\) for \(s>4m_1^2\)
on the physical upper boundary of the first sheet, in the terminology of
Definition~\ref{def:stable-channel-sheet-convention}. Elastic unitarity gives
\begin{equation}
  |S_\ell(s)|=1 .
  \label{eq:resonance-elastic-unitarity}
\end{equation}
With inelastic channels included on the same Hilbert space, the elastic
matrix element instead obeys \(|S_\ell(s)|\le1\). Thus the exact physical-axis
partial wave is finite. The divergence of
\(\mathcal M_s^{\mathrm{tree}}\) at \(s=m_2^2\) records that the perturbative
description has expanded around a stable reference line in a region where the
state couples to an open decay channel. The exact object is a resonance: a
nearby pole of the continuation of \(S_\ell(s)\), or equivalently of
\(a_\ell(s)\), through the threshold cut. On the physical sheet the
same channel is represented by the continuum part of the spectral measure;
an isolated real mass atom occurs only for a stable one-particle state.

\begin{proposition}[Physical-axis unitarity excludes the resonance pole]
\label{prop:physical-axis-unitarity-excludes-resonance-pole}
Let \(I\subset(4m_1^2,\infty)\) be an elastic interval away from the threshold
endpoint, and assume that the exact stable-channel partial-wave boundary value
exists on \(I\).  If \(S_\ell(s)\) is unitary on \(I\), or if it is the elastic
matrix element of a larger unitary channel matrix, then
\[
  |a_\ell(s)|
  \le
  \frac{1}{\rho_1(s)}
  \qquad (s\in I).
\]
Consequently a resonance associated with this channel is not a pole of the
physical upper-edge boundary value \(a_\ell(s+\ii0)\).  It is a pole of the
analytic continuation through the channel cut.
\end{proposition}

\begin{proof}
Equation~\eqref{eq:resonance-elastic-s-matrix-normalization} gives
\[
  a_\ell(s)=\frac{S_\ell(s)-1}{2\ii\rho_1(s)} .
\]
On \(I\), \(\rho_1(s)>0\).  If \(S_\ell\) is elastic unitary, then
\(|S_\ell|=1\), while if it is the elastic component of a larger unitary
matrix, then \(|S_\ell|\le1\).  In both cases
\[
  |a_\ell(s)|
  \le
  \frac{|S_\ell(s)|+1}{2\rho_1(s)}
  \le
  \frac{1}{\rho_1(s)} .
\]
The bound is finite on compact subintervals away from threshold.  Hence a
simple divergence on the physical boundary is incompatible with exact
stable-channel unitarity; the tree pole is a perturbative signal of a nearby
singularity on an analytically continued sheet, not a physical-axis pole.
\end{proof}

\section{Outgoing Poles in a One-Dimensional Model}
\label{sec:resonance-one-dimensional-outgoing-poles}

This section is the detailed one-dimensional model promised in
Section~\ref{sec:bound-state-one-dimensional-orientation}.  It extends the
bound-state pole criterion to outgoing resonance poles; the underlying
Hilbert-space interpretation of generalized scattering states is the one fixed
in Section~\ref{sec:rigged-hilbert-continuous-spectrum}.  Let
\[
  H=-\frac{1}{2\mu}\frac{\dd^2}{\dd x^2}+V(x),
\]
where \(V\) is real and short range. For \(x\to+\infty\), use the outgoing
solution
\[
  \psi_k^{\mathrm{out}}(x,t)
  \sim
  S(k)^{-1}\ee^{-\ii kx-\ii Et}
  +
  \ee^{\ii kx-\ii Et},
  \qquad
  E=\frac{k^2}{2\mu}.
\]

\begin{definition}[Outgoing one-dimensional resonance pole]
\label{def:outgoing-one-dimensional-resonance-pole}
For the one-dimensional short-range Hamiltonian above, an outgoing resonance
pole is a pole \(k_\ast\) of the meromorphic continuation of \(S(k)\) with
\[
  \operatorname{Re}k_\ast>0,
  \qquad
  \operatorname{Im}k_\ast<0,
\]
for which the corresponding continued solution satisfies the outgoing
boundary condition at spatial infinity.  Its complex energy is
\[
  E_\ast=\frac{k_\ast^2}{2\mu}.
\]
\end{definition}

\begin{proposition}[Outgoing pole gives decay in time and growth in space]
\label{prop:outgoing-pole-time-decay-space-growth}
If \(S(k)\) has a pole at \(k=k_\ast\), then the incoming coefficient
vanishes at that analytically continued value and the asymptotic solution is
purely outgoing. Write
\[
  E_\ast=\frac{k_\ast^2}{2\mu}=\omega-\ii\gamma,
  \qquad
  \gamma>0.
\]
Then
\[
  \psi_{k_\ast}^{\mathrm{out}}(x,t)
  \sim
  \ee^{\ii k_\ast x-\ii\omega t-\gamma t}.
\]
The negative imaginary part of \(E_\ast\) is the exponential decay rate of
the time dependence. The same wave grows in space along the outgoing
direction. This spatial growth is forced by probability conservation for a
state with only outgoing flux: a uniformly decaying packet cannot also be a
normalizable stationary eigenvector of a self-adjoint Hamiltonian. The pole
therefore defines a singular outgoing functional, a resonance or Gamow
functional, associated with the analytically continued scattering function.
\end{proposition}

\begin{proof}
At a simple pole of \(S(k)\), the coefficient \(S(k)^{-1}\) of the incoming
wave in the displayed outgoing solution vanishes.  The continued asymptotic
solution is therefore proportional to
\(\exp(\ii k_\ast x-\ii E_\ast t)\).  Write
\[
  k_\ast=a-\ii b,\qquad a>0,\quad b>0 .
\]
Then
\[
  E_\ast=\frac{a^2-b^2}{2\mu}-\ii\frac{ab}{\mu}
  =
  \omega-\ii\gamma,
  \qquad
  \gamma=\frac{ab}{\mu}>0.
\]
The time factor is \(\exp(-\ii\omega t)\exp(-\gamma t)\), while the spatial
factor for \(x\to+\infty\) contains \(\exp(bx)\).  Thus the same analytic
continuation that gives decay for \(t>0\) gives exponential growth in the
outgoing spatial direction.  Such a function is not in \(L^2(\R)\), so it is
not an eigenvector of the self-adjoint Hamiltonian on the Hilbert space.
\end{proof}

The spatial growth and temporal decay described in the proof are summarized in
Figure~\ref{fig:resonance-k-plane-outgoing-growth}.
\begin{figure}[!ht]
\centering
\begin{tikzpicture}[>=Stealth,scale=0.9]
  \begin{scope}
    \draw[->,line width=0.7pt] (-0.15,0)--(5.65,0) node[right] {\(x\)};
    \draw[->,line width=0.7pt] (0,-1.65)--(0,2.15) node[above] {\(V(x)\)};
    \draw[blue!70!black,line width=1.15pt]
      (0.25,1.75) .. controls (0.55,-0.95) and (1.08,-1.5) .. (1.48,-0.18)
      .. controls (1.85,0.92) and (2.18,0.78) .. (2.55,0.12)
      .. controls (3.1,-0.65) and (3.85,0.2) .. (5.2,0.05);
    \draw[orange!90!black,line width=1.05pt,dash pattern=on 5pt off 2.5pt]
      (1.05,0.36) .. controls (1.3,0.75) and (1.58,0.75) .. (1.82,0.28)
      .. controls (2.02,-0.12) and (2.32,0.04) .. (2.66,0.36)
      .. controls (3.08,0.7) and (3.72,0.28) .. (4.92,-0.18);
    \foreach \x in {1.35,1.82,2.42,3.42,4.25}
      \draw[brown!70!black,->,line width=0.8pt] (\x,0.18)--++(0,-0.48);
    \draw[red!80!black,->,line width=1.05pt] (4.05,0.92)--(5.05,0.92);
    \node[red!80!black,align=left,anchor=west,font=\small] at (5.16,0.92)
      {outgoing\\flux};
    \node[orange!90!black,align=center,font=\small] at (1.08,1.05)
      {temporarily\\localized};
  \end{scope}
  \begin{scope}[xshift=8.0cm]
    \draw[->,line width=0.7pt] (-1.65,0)--(1.95,0)
      node[right] {\(\operatorname{Re}k\)};
    \draw[->,line width=0.7pt] (0,-1.55)--(0,1.9)
      node[above] {\(\operatorname{Im}k\)};
    \draw[blue!70!black,line width=1.1pt,->] (0,0)--(1.45,0);
    \fill[red!75!black] (0,0.85) circle (2.3pt);
    \node[red!75!black,anchor=west,font=\small] at (0.18,0.86)
      {bound};
    \fill[orange!90!black] (0.84,-0.70) rectangle (1.00,-0.54);
    \node[orange!90!black,anchor=west,font=\small] at (1.08,-0.62)
      {\(k_\ast\)};
    \draw[orange!90!black,->,line width=0.8pt] (0.55,-0.18)--(0.86,-0.55);
    \node at (0,-1.95) {\(E=k^2/(2\mu)\)};
  \end{scope}
\end{tikzpicture}
\caption{Outgoing Gamow behavior and the \(k\)-plane location of poles in the
one-dimensional resonance model.  The dashed packet in the left panel marks
temporary localization and the arrow marks outward flux.  In the right panel a
circular marker denotes a bound-state pole on the positive imaginary \(k\)-axis,
while a square marker denotes a resonance pole with
\(\operatorname{Im} k_\ast<0\).}
\label{fig:resonance-k-plane-outgoing-growth}
\end{figure}

Because \(E=k^2/(2\mu)\), the energy plane is a two-sheeted cover of the
\(k\)-plane. Bound-state poles lie on the positive imaginary \(k\)-axis and
map to negative real energies on the physical sheet. Resonance poles with
\(\operatorname{Re}k_\ast>0\) and \(\operatorname{Im}k_\ast<0\) map to
complex energies below the positive real axis on the sheet reached by
continuing through the scattering cut.  The continuation through the cut is
shown explicitly in Figure~\ref{fig:resonance-second-sheet-energy-plane}.

\begin{figure}[!ht]
\centering
\begin{tikzpicture}[>=Stealth,scale=0.92]
  \draw[->,line width=0.7pt] (-2.25,0)--(4.45,0)
    node[right] {\(\operatorname{Re}E\)};
  \draw[->,line width=0.7pt] (0,-1.65)--(0,2.05)
    node[above] {\(\operatorname{Im}E\)};
  \draw[blue!70!black,line width=1.05pt,->] (0,0)--(3.85,0);
  \node[blue!70!black,below,font=\small] at (1.75,-0.14)
    {\(\mathbb R_+\) cut};
  \fill[red!75!black] (-1.18,0) circle (2.4pt);
  \fill[red!75!black] (-0.62,0) circle (2.4pt);
  \node[red!75!black,align=center,font=\small] at (-1.22,0.72)
    {bound\\states};
  \fill[orange!90!black] (2.35,-0.78) circle (2.5pt);
  \node[orange!90!black,anchor=west,font=\small] at (2.52,-0.78)
    {resonance pole};
  \draw[orange!90!black,dashed,line width=0.75pt] (2.35,-0.78)--(2.35,0);
  \draw[purple,->,line width=0.95pt]
    (0.78,0.72) .. controls (1.42,1.28) and (2.52,1.18) .. (3.14,0.56);
  \draw[gray!65!black,dashed,line width=0.75pt]
    (3.02,0.18)--(3.02,-0.28);
  \node[gray!65!black,anchor=west,font=\scriptsize] at (3.10,0.34)
    {through cut};
  \draw[green!55!black,->,line width=0.95pt]
    (3.0,-0.32) .. controls (2.86,-0.72) and (2.62,-0.92) .. (2.40,-0.80);
  \node[purple,align=center,font=\small] at (1.55,1.42)
    {first sheet\\upper lip};
  \node[green!55!black,align=center,font=\small] at (0.72,-1.18)
    {second sheet\\lower lip};
  \fill (0,0) circle (1.8pt) node[below left] {\(0\)};
\end{tikzpicture}
\caption{Energy-plane branch structure for a resonance pole.  The first sheet
is reached from the upper edge of the positive real cut; crossing the cut gives
the second-sheet lower-edge branch on which the resonance pole appears.}
\label{fig:resonance-second-sheet-energy-plane}
\end{figure}

\section{The Dressed Propagator}

Return to the relativistic exchange model.
\begin{definition}[Dressed propagator and self-energy convention]
\label{def:dressed-propagator-self-energy-convention}
The exact time-ordered two-point function of \(\phi_2\) in momentum space is
\[
  D_2(k)
  =
  \int \dd^4x\,\ee^{-\ii k\cdot x}
  \bra{\vac}T\phi_2(x)\phi_2(0)\ket{\vac} .
\]
Let \(\Sigma(k)\) be the sum of amputated one-particle-irreducible two-point
insertions for the \(\phi_2\) line, with the convention
\[
  D_2(k)=\frac{-\ii}{k^2+m_{2,\Lambda}^2-\ii0-\Sigma(k)} .
\]
This is the geometric series obtained by inserting 1PI two-point kernels into
the regulated reference propagator.
\end{definition}

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.84,
  two/.style={dash pattern=on 3pt off 2pt,line width=0.85pt},
  one/.style={line width=0.85pt},
  blob/.style={draw,circle,minimum size=0.58cm,inner sep=0pt}]
  \node at (-5.25,0) {\(D_2=\)};
  \draw[two] (-4.45,0)--(-3.35,0);
  \node at (-2.9,0) {\(+\)};
  \draw[two] (-2.35,0)--(-1.62,0);
  \node[blob] at (-1.25,0) {\scriptsize 1PI};
  \draw[two] (-0.88,0)--(-0.15,0);
  \node at (0.35,0) {\(+\)};
  \draw[two] (0.9,0)--(1.45,0);
  \node[blob] at (1.82,0) {\scriptsize 1PI};
  \draw[two] (2.18,0)--(2.72,0);
  \node[blob] at (3.1,0) {\scriptsize 1PI};
  \draw[two] (3.47,0)--(4.02,0);
  \node at (4.55,0) {\(+\cdots\)};
  \begin{scope}[yshift=-2.1cm]
    \node at (-3.75,0) {\(\ii\Sigma_1(k)=\)};
    \draw[two] (-2.45,0)--(-1.75,0);
    \draw[two] (-0.55,0)--(0.15,0);
    \draw[one] (-1.15,0) circle (0.6);
    \node[blue!70!black] at (-1.12,0.88) {\(\ell\)};
    \node[blue!70!black] at (-1.08,-0.88) {\(k-\ell\)};
    \node at (0.82,0) {\(+O(g^4)\)};
  \end{scope}
\end{tikzpicture}
\end{center}

At one loop,
\[
\begin{aligned}
  \ii\Sigma_1(k)
  &=
  \frac{(-\ii g)^2}{2}
  \int\frac{\dd^D\ell}{(2\pi)^D}\,
  \frac{-\ii}{\ell^2+m_1^2-\ii0}\,
  \frac{-\ii}{(k-\ell)^2+m_1^2-\ii0}.
\end{aligned}
\]
The factor \(1/2\) is the symmetry factor for the two identical \(\phi_1\)
lines in the loop. Combining denominators gives
\[
  \ii\Sigma_1(k)
  =
  \frac{g^2}{2}
  \int_0^1\dd x
  \int\frac{\dd^D\ell}{(2\pi)^D}\,
  \frac{1}{
    \left[\ell^2(1-x)+(k-\ell)^2x+m_1^2-\ii0\right]^2}.
\]
Shift \(\ell^\mu\mapsto\ell^\mu+xk^\mu\), then rotate the shifted energy
contour according to the Feynman prescription. The Euclidean expression is
\[
  \ii\Sigma_1(k)
  =
  \ii\frac{g^2}{2}
  \int_0^1\dd x
  \int\frac{\dd^D\ell_E}{(2\pi)^D}\,
  \frac{1}{
    \left[\ell_E^2+k^2x(1-x)+m_1^2\right]^2}.
\]
The rotation is a contour statement, not a change of notation: for the
shifted loop variable the pole displacements prescribed by \(-\ii0\) leave a
sector through which the \(\ell^0\)-contour can be moved to the Euclidean
axis without crossing singularities.  After the contour has been rotated, the
Euclidean denominator is an ordinary real denominator.  Its boundary value on a
timelike cut is obtained by analytically continuing the resulting function with
the original Feynman prescription; it is not encoded by leaving a residual
\(-\ii0\) in the Euclidean integrand.  When the answer is later written as a
Lorentzian boundary value, a notation such as
\(k^2x(1-x)+m_1^2-\ii0\), or
\(1-sx(1-x)/m_1^2-\ii0\) with \(k^2=-s\), refers to the boundary-value
choice of the external invariant \(k^2\) only.  It is not a pole prescription
for the already rotated Euclidean loop variable \(\ell_E\).

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.9]
  \draw[->,line width=0.7pt] (-2.8,0)--(2.8,0)
    node[right] {\(\operatorname{Re}\ell^0\)};
  \draw[->,line width=0.7pt] (0,-2.0)--(0,2.2)
    node[above] {\(\operatorname{Im}\ell^0\)};
  \draw[purple,line width=1.0pt,->] (-2.45,0)--(2.25,0);
  \draw[green!55!black,line width=1.0pt,->] (0,-1.7)--(0,1.75);
  \fill[red!75!black] (-1.12,0.35) circle (2.3pt);
  \fill[red!75!black] (1.15,-0.35) circle (2.3pt);
  \node[red!75!black,font=\small] at (-1.62,0.82)
    {Feynman poles};
  \draw[orange!90!black,->,line width=1.0pt]
    (1.6,0.18) .. controls (1.25,0.92) and (0.62,1.28) .. (0.08,1.34);
  \node[purple,font=\small,below] at (-1.62,-0.12) {Lorentzian};
  \node[green!55!black,font=\small,anchor=west] at (0.18,1.58)
    {Euclidean};
\end{tikzpicture}
\end{center}

In four dimensions, with a Euclidean cutoff \(|\ell_E|<\Lambda\) and with
\(k\) in the domain reached by the contour rotation,
\[
  \Sigma_1(k)
  =
  \frac{g^2}{32\pi^2}
  \int_0^1\dd x
  \left[
    \log
    \frac{\Lambda^2}{k^2x(1-x)+m_1^2}
    -1+O(\Lambda^{-2})
  \right].
\]
The divergent part is independent of \(k\). Introduce
\[
  m_{2,\Lambda}^2=(m_2^R)^2+\delta m_2^2,
  \qquad
  \Sigma_1(k)=\Sigma_1^R(k)+\delta m_2^2 ,
\]
where the counterterm cancels the cutoff-dependent constant. One convenient
subtraction is
\[
  \Sigma_1^R(k)
  =
  -\frac{g^2}{32\pi^2}
  \int_0^1\dd x\,
  \log
  \frac{k^2x(1-x)+m_1^2}{m_1^2}.
\]
This is first an analytic function on the Euclidean side of the cut.  The
physical boundary value is specified only after the external invariant is
continued to a timelike value.

\section{The Threshold Cut and the Cut Width}

For the \(s\)-channel exchange, \(k=K=k_1+k_2\) and \(K^2=-s\). The dressed
contribution is
\[
  \mathcal M_s(s)
  =
  \frac{g^2}{F(s)},
  \qquad
  F(s)=
  -s+(m_2^R)^2-\ii0-\Sigma_1^R(K).
\]
Equivalently,
\[
  F(s)
  =
  -s+(m_2^R)^2-\ii0
  +
  \frac{g^2}{32\pi^2}
  \int_0^1\dd x\,
  \log\left[
    1-\frac{s\,x(1-x)}{m_1^2}-\ii0
  \right].
\]
Since
\[
  0\le x(1-x)\le\frac14,
\]
the logarithm develops a discontinuity precisely for
\[
  s\ge4m_1^2 .
\]
For real \(s>4m_1^2\), define
\[
  \rho_1(s)=\sqrt{1-\frac{4m_1^2}{s}},
  \qquad
  x_\pm=\frac12\left(1\pm\rho_1(s)\right).
\]
The argument of the logarithm is negative for \(x\in[x_-,x_+]\). On the
physical upper boundary the prescription gives
\[
  \log(-|a|-\ii0)=\log|a|-\ii\pi ,
\]
and therefore
\[
  \operatorname{Im}\Sigma_1^R(K)\big|_{s+\ii0}
  =
  \frac{g^2}{32\pi}\rho_1(s).
\]
This boundary value is fixed by the same Wick-rotation prescription that
relates the Euclidean Green function to the Lorentzian time-ordered Green
function. For real positive energy above the two-particle threshold,
\(k^0\) is reached by analytic continuation from the imaginary \(k^0\)-axis
to the positive real axis through the upper half-plane.

\begin{proposition}[Threshold discontinuity of the one-loop self-energy]
\label{prop:threshold-discontinuity-one-loop-self-energy}
For \(s>4m_1^2\), with the physical upper-edge boundary value specified
above,
\[
  \operatorname{Im}\Sigma_1^R(K)\big|_{K^2=-s+\ii0}
  =
  \frac{g^2}{32\pi}\rho_1(s),
  \qquad
  \rho_1(s)=\sqrt{1-\frac{4m_1^2}{s}} .
\]
Equivalently, the imaginary part of the dressed inverse denominator on the
first sheet is
\[
  \operatorname{Im}F_{\rm I}(s+\ii0)
  =
  -\frac{g^2}{32\pi}\rho_1(s)+O(0^+).
\]
\end{proposition}

\begin{proof}
The logarithm in \(F(s)\) is
\[
  \log\left[1-\frac{s\,x(1-x)}{m_1^2}-\ii0\right].
\]
Its argument is negative precisely when \(x\in[x_-,x_+]\), where
\[
  x_\pm=\frac12(1\pm\rho_1(s)).
\]
The interval length is \(x_+-x_-=\rho_1(s)\).  On the physical upper edge,
\[
  \log(-|a|-\ii0)=\log|a|-\ii\pi .
\]
Since
\[
  \Sigma_1^R(K)
  =
  -\frac{g^2}{32\pi^2}
  \int_0^1\dd x\,
  \log\left[
    1-\frac{s\,x(1-x)}{m_1^2}-\ii0
  \right],
\]
the imaginary part is
\[
  -\frac{g^2}{32\pi^2}(-\pi)\,(x_+-x_-)
  =
  \frac{g^2}{32\pi}\rho_1(s).
\]
The denominator \(F_{\rm I}=m_R^2-s-\ii0-\Sigma_1^R\) therefore has the
opposite imaginary part, up to the infinitesimal boundary prescription.
\end{proof}

Figure~\ref{fig:self-energy-cut-continuation} records the two analytic
continuations used in the self-energy computation: the \(k^0\) contour
deformation and the corresponding discontinuity across the \(s\)-channel cut.
\begin{figure}[!ht]
\centering
\begin{tikzpicture}[>=Stealth,scale=0.92]
  \begin{scope}
    \draw[->,line width=0.7pt] (-2.55,0)--(3.0,0)
      node[right] {\(\operatorname{Re}k^0\)};
    \draw[->,line width=0.7pt] (0,-2.0)--(0,2.25)
      node[above] {\(\operatorname{Im}k^0\)};
    \draw[green!55!black,line width=1.0pt] (0,-1.72)--(0,1.85);
    \draw[red!75!black,decorate,
      decoration={snake,amplitude=0.55mm,segment length=2.7mm},
      line width=0.95pt] (1.05,0)--(2.75,0);
    \draw[red!75!black,decorate,
      decoration={snake,amplitude=0.55mm,segment length=2.7mm},
      line width=0.95pt] (-2.35,0)--(-1.05,0);
    \draw[orange!90!black,->,line width=1.05pt]
      (0.22,1.35) .. controls (1.18,1.2) and (1.88,0.72) .. (2.12,0.12);
    \node[orange!90!black,align=center,font=\small] at (1.65,1.48)
      {continue\\from above};
  \end{scope}
  \begin{scope}[xshift=6.6cm]
    \draw[->,line width=0.7pt] (-0.4,0)--(5.05,0) node[right] {\(s\)};
    \draw[->,line width=0.7pt] (0,-1.75)--(0,2.0)
      node[above] {\(\operatorname{Im}s\)};
    \fill (1.28,0) circle (2pt) node[below] {\(0\)};
    \fill (2.62,0) circle (2pt) node[below] {\(4m_1^2\)};
    \draw[red!75!black,decorate,
      decoration={snake,amplitude=0.55mm,segment length=2.7mm},
      line width=1.05pt] (2.62,0)--(4.78,0);
    \node[purple,align=center,font=\small] at (3.72,1.1)
      {\(\operatorname{Im}\log=-\pi\)\\upper edge};
    \node[cyan!70!black,align=center,font=\small] at (3.72,-1.08)
      {\(\operatorname{Im}\log=+\pi\)\\lower edge};
    \draw[purple,->,line width=0.9pt] (3.2,0.78)--(3.2,0.12);
    \draw[cyan!70!black,->,line width=0.9pt] (3.2,-0.78)--(3.2,-0.12);
  \end{scope}
\end{tikzpicture}
\caption{Cut continuation for the one-loop self-energy.  The left panel shows
the \(k^0\) continuation from the upper edge past the on-shell singularities.
The right panel records the sign convention for the logarithm on the two sides
of the \(s\)-channel cut beginning at \(4m_1^2\).}
\label{fig:self-energy-cut-continuation}
\end{figure}

Choose a real resonance mass parameter \(M_R\) by the condition that the real
part of the denominator vanishes at \(s=M_R^2\),
\[
  \operatorname{Re}F(M_R^2+\ii0)=0 .
\]
At weak coupling, and to leading order in the variation of the real part of
\(\Sigma_1^R\), the amplitude near this point is
\[
  \mathcal M_s(s)
  \simeq
  \frac{g^2}{M_R^2-s-\ii M_R\Gamma_{\rm cut}(M_R)},
\]
where
\[
  M_R\Gamma_{\rm cut}(M_R)
  =
  \frac{g^2}{32\pi}
  \sqrt{1-\frac{4m_1^2}{M_R^2}}
  +O(g^4).
\]
The subscript ``cut'' is deliberate.  The number above is extracted from the
absorptive part of the two-point function on the physical cut.  It is the
quantity computed by the inclusive cut, or Fermi golden rule, calculation at
the real energy \(M_R\).  The underlying discontinuity is the two-point
specialization of Hilbert-space unitarity; the line-replacement calculation is
its perturbative expansion, in the sense explained in
Remark~\ref{rem:unitarity-discontinuity-versus-cutkosky}.  It is not, by
itself, a definition of every resonance width.

\section{Width Parameters and Their Comparison}

There are several operationally different quantities that are often denoted by
the same letter \(\Gamma\).  In a narrow isolated resonance they agree to the
order at which the local Breit--Wigner approximation is meaningful, but their
definitions are distinct.

\begin{definition}[Width coordinates for an isolated resonance]
\label{def:resonance-width-coordinates}
Let a stable-channel amplitude have an isolated simple resonance pole
\(s_\ast\) on a specified nonphysical sheet, and assume that a real energy
window containing the corresponding line shape has been fixed.  The following
four quantities are different coordinates on different data.

\begin{enumerate}
\item The cut width \(\Gamma_{\rm cut}(M)\) is defined at a real energy \(M\)
by the inclusive transition rate obtained from the discontinuity of the
self-energy at that real energy.  For a scalar resonance coupled to stable
final channels \(X\), the perturbative expression has the form
\[
  \Gamma_{\rm cut}(M)
  =
  \frac{1}{2M}
  \sum_X
  \int \dd\Phi_X(P)\,
  \left|\mathcal A(R(P)\to X)\right|^2 ,
  \qquad P^2=-M^2,\quad P^0>0,
\]
with the symmetry factors included in the invariant phase-space measure
\(\dd\Phi_X(P)\) and in the amplitude \(\mathcal A\).  This formula should be
read as a weak-coupling construction using the stable state \(R\) of the
decoupled or lower-order theory as an interpolating state.  Once the channel is
open, the exact theory has stable asymptotic states in the final channels, not
an exact external one-resonance state.

\item The line-shape width \(\Gamma_{\rm line}\) is an observable extracted
from a specified physical inclusive cross section or event distribution.  After a
specified smooth background subtraction and a specified resolution function,
one may define it as the full width at half maximum in the energy variable
\(E=\sqrt{s}\) of the resonant contribution.  This definition depends on the
chosen process, on the observable used to form the inclusive distribution, and
on how nonresonant background is separated from the pole term.  It is therefore
not an invariant analytic datum for a broad resonance.

\item The Gamow time width, or complex-energy pole width, is defined from the
second-sheet pole after choosing an energy square root.  If \(s_\ast\) is a
simple resonance pole and the square-root branch is chosen so that
\(\operatorname{Re}\sqrt{s_\ast}>0\) near the resonance, write
\[
  E_\ast=\sqrt{s_\ast}
  =
  M_{\rm pole}-\frac{\ii}{2}\Gamma_{\rm pole},
  \qquad M_{\rm pole}>0,\quad \Gamma_{\rm pole}>0.
\]
Equivalently,
\[
  s_\ast
  =
  M_{\rm pole}^2
  -\ii M_{\rm pole}\Gamma_{\rm pole}
  -\frac14\Gamma_{\rm pole}^2 .
\]
This definition is independent of a particular production process once the
pole and sheet have been specified.  The associated outgoing Gamow functional
evolves as
\[
  \ee^{-\ii E_\ast t}
  =
  \ee^{-\ii M_{\rm pole}t}\,
  \ee^{-\Gamma_{\rm pole}t/2},
\]
so its probability norm, in the formal Gamow description, decays as
\(\ee^{-\Gamma_{\rm pole}t}\).  This is the time-decay width of the pole
functional.  The statement concerns the pole contribution in the rigged-Hilbert
analytic continuation; an exact normalizable state in a self-adjoint
Hamiltonian does not have a purely exponential survival law for all times.

\item The \(s\)-plane pole-imaginary width is defined by
\[
  M_s=\sqrt{\operatorname{Re}s_\ast},
  \qquad
  \Gamma_s=-\frac{\operatorname{Im}s_\ast}{M_s}.
\]
This convention uses the imaginary part of the pole in the \(s\)-plane rather
than the imaginary part of the energy \(E=\sqrt{s}\).  It agrees with
\(\Gamma_{\rm pole}\) only up to the kinematic conversion
\[
  \Gamma_s
  =
  \Gamma_{\rm pole}
  \left[
    1+O\!\left(\frac{\Gamma_{\rm pole}^2}{M_{\rm pole}^2}\right)
  \right].
\]
\end{enumerate}
\end{definition}

\begin{controlledapproximation}[Narrow-resonance comparison of widths]
Suppose that, in the energy window
\(|E-M_{\rm pole}|=O(\Gamma_{\rm pole})\), the continued stable-channel
amplitude is dominated by a single simple pole, the residue and the
nonresonant background are holomorphic and vary only on a scale of order
\(M_{\rm pole}\), and the pole is separated from thresholds and other
singularities by a squared-mass distance much larger than
\(M_{\rm pole}\Gamma_{\rm pole}\).  Then the ideal resonant line shape
obtained from the pole term, with the smooth residue replaced by its value at
the pole and with the smooth background removed, has
\[
  \Gamma_{\rm line}
  =
  \Gamma_{\rm pole}
  \left[
    1+O\!\left(\frac{\Gamma_{\rm pole}^2}{M_{\rm pole}^2}\right)
  \right],
\]
as follows.  Keeping only the pole term and writing
\(E=M_{\rm pole}+x\),
\[
  |s_\ast-E^2|^2
  =
  \left(
    2M_{\rm pole}x+x^2+\frac14\Gamma_{\rm pole}^2
  \right)^2
  +
  M_{\rm pole}^2\Gamma_{\rm pole}^2 .
\]
The half-maximum points of the pole term, measured in the energy variable, are
therefore
\[
  x_\pm
  =
  \pm\frac12\Gamma_{\rm pole}
  -\frac{1}{4M_{\rm pole}}\Gamma_{\rm pole}^2
  +O\!\left(\frac{\Gamma_{\rm pole}^3}{M_{\rm pole}^2}\right),
\]
so \(x_+-x_-=\Gamma_{\rm pole}[1+O(\Gamma_{\rm pole}^2/M_{\rm pole}^2)]\).
The common shift of the two half-maximum points records that the peak in
\(E\) is not centered exactly at \(M_{\rm pole}\) once the quadratic relation
\(s=E^2\) is retained.  If the residue, phase space, detector resolution, or
subtracted background varies appreciably over the same energy interval, the
measured line-shape width receives additional process-dependent corrections.
That is precisely why \(\Gamma_{\rm line}\) is not an invariant analytic datum
for a broad resonance.

The cut calculation at the real resonance energy gives
\[
  M_{\rm pole}\Gamma_{\rm pole}
  =
  M_R\Gamma_{\rm cut}(M_R)
  +O(g^4)
\]
in the weak-coupling scalar model above.  Indeed, on the second sheet the
discontinuity computation gives
\[
  F_{\rm II}(M_R^2)
  =
  -\ii M_R\Gamma_{\rm cut}(M_R)+O(g^4),
  \qquad
  F_{\rm II}'(M_R^2)=-1+O(g^2).
\]
Solving \(F_{\rm II}(s_\ast)=0\) by one Newton step therefore gives
\[
  s_\ast-M_R^2
  =
  -\ii M_R\Gamma_{\rm cut}(M_R)+O(g^4).
\]
Thus the common symbol \(\Gamma_R\) is justified only after these hypotheses
and approximations have been stated.  For a broad resonance, for an
overlapping pair of poles, or near a threshold branch point,
\(\Gamma_{\rm line}\), \(\Gamma_{\rm cut}\), \(\Gamma_{\rm pole}\), and
\(\Gamma_s\) need not coincide.
\end{controlledapproximation}

\section{Elastic Phase Motion}

The preceding weak-coupling discussion defines a single-channel elastic
resonance model after three choices have been made: retain the resonant
\(s\)-channel exchange, discard smooth background and crossed-channel pieces,
and replace the absorptive part by the width function
\[
  W(s)=\frac{g^2}{32\pi}\rho_1(s)
\]
on the elastic real interval.  These choices are the approximation.  Once they
have been made, the corresponding Breit--Wigner elastic \(S\)-matrix element is
a definite rational function and its elastic unitarity is exact in the model.

Only the \(\ell=0\) partial wave carries the retained scalar exchange.  Since
the resonant \(s\)-channel term is independent of \(z\),
Equation~\eqref{eq:resonance-partial-wave-expansion} gives
\[
  a_0^{(s)}(s)=\frac{\mathcal M_s(s)}{32\pi},
  \qquad
  S_0(s)-1=2\ii\rho_1(s)a_0^{(s)}(s)
  =
  \frac{\ii\rho_1(s)}{16\pi}\mathcal M_s(s).
\]
Equivalently,
\[
  \mathcal M_s(s)
  =
  -16\pi\ii\sqrt{\frac{s}{s-4m_1^2}}\,
  \bigl(S_0(s)-1\bigr),
\]
for the angle-independent elastic partial wave in this normalization.  The
Breit--Wigner elastic approximant is therefore
\[
  S_0^{\rm BW}(s)
  :=
  1+
  \frac{\ii\rho_1(s)}{16\pi}
  \frac{g^2}{M_R^2-s-\ii W(s)}
  =
  \frac{
    M_R^2-s+\ii W(s)
  }{
    M_R^2-s-\ii W(s)
  },
  \qquad
  W(s)
  =
  \frac{g^2}{32\pi}\rho_1(s).
\]

\begin{proposition}[Elastic unitarity of the Breit--Wigner factor]
\label{prop:breit-wigner-elastic-unitarity}
On a real interval where only the elastic \(\phi_1\phi_1\) channel is retained
and \(W(s)\in\R\), the model \(S\)-matrix element
\[
  S_0^{\rm BW}(s)=
  \frac{M_R^2-s+\ii W(s)}{M_R^2-s-\ii W(s)}
\]
obeys
\[
  |S_0^{\rm BW}(s)|=1
\]
wherever the denominator is nonzero.
\end{proposition}

\begin{proof}
For real \(s\), the number \(A(s)=M_R^2-s\) is real and, by hypothesis,
\(W(s)\) is real.  Hence \(A(s)+\ii W(s)\) is the complex conjugate of
\(A(s)-\ii W(s)\).  Therefore
\[
  |S_0^{\rm BW}(s)|^2
  =
  \frac{(A+\ii W)(A-\ii W)}{(A-\ii W)(A+\ii W)}
  =
  1 .
\]
\end{proof}

The physical partial wave differs from \(S_0^{\rm BW}\) by the omitted
background, crossed-channel, inelastic, and higher-order corrections; these
are corrections to the model, not violations of the algebraic unitarity of the
Breit--Wigner factor itself.  The resonance is expressed as rapid phase
motion. If
\[
  S_0^{\rm BW}(s)=\ee^{2\ii\delta_0^{\rm BW}(s)},
\]
then the phase shift \(\delta_0^{\rm BW}\) changes rapidly through the point
where \(\operatorname{Re}F_{\rm II}(s)\) vanishes in the same local
second-sheet approximation.

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.95]
  \begin{scope}
    \draw[->,line width=0.75pt] (-2.0,0)--(2.15,0)
      node[right] {\(\operatorname{Re}S_0^{\rm BW}\)};
    \draw[->,line width=0.75pt] (0,-2.0)--(0,2.15)
      node[above] {\(\operatorname{Im}S_0^{\rm BW}\)};
    \draw[green!55!black,dashed,line width=0.9pt] (0,0) circle (1.42);
    \draw[blue!70!black,line width=1.25pt]
      (1.38,-0.08) arc[start angle=-3,end angle=345,radius=1.42];
    \draw[blue!70!black,->,line width=1.0pt]
      (0.2,-1.4) arc[start angle=-82,end angle=-20,radius=1.42];
    \node[green!55!black,anchor=west,font=\small] at (1.78,1.0)
      {\(|S_0^{\rm BW}|=1\)};
  \end{scope}
  \begin{scope}[xshift=6.4cm]
    \draw[->,line width=0.75pt] (-0.25,0)--(5.2,0) node[right] {\(s\)};
    \draw[->,line width=0.75pt] (0,-1.45)--(0,1.9)
      node[above] {\(\operatorname{Im}s\)};
    \draw[red!75!black,decorate,
      decoration={snake,amplitude=0.55mm,segment length=2.7mm},
      line width=1.05pt] (2.15,0)--(4.9,0);
    \node[red!75!black,below] at (2.15,-0.12) {\(4m_1^2\)};
    \fill[orange!90!black] (3.35,-0.72) circle (2.4pt);
    \node[orange!90!black,anchor=west,font=\small] at (3.52,-0.72)
      {second-sheet pole};
    \draw[orange!90!black,dashed,line width=0.75pt] (3.35,-0.72)--(3.35,0);
    \node[orange!90!black,above,font=\small] at (3.35,0.2) {\(M_R^2\)};
  \end{scope}
\end{tikzpicture}
\end{center}

\section{The Second Sheet}

The sheet convention was fixed in
Definition~\ref{def:stable-channel-sheet-convention}.  Here it is made
explicit for the above-threshold exchange model.  The terminology applies to
connected stable-channel scattering kernels.

\begin{definition}[Adjacent second sheet in a single stable channel]
\label{def:adjacent-second-sheet-single-channel}
After the identity term and the overall momentum-conservation delta function
have been removed, let
\[
  \mathcal M_{ab}^{\rm I}(s)
\]
denote a connected invariant amplitude or partial-wave coordinate between
stable channels \(a\) and \(b\) on the first sheet.  Let
\(I\subset(4m_1^2,\infty)\) be an open interval on which the only cut being
crossed is the \(\phi_1\phi_1\) threshold cut and on which the upper and lower
first-sheet boundary values exist as distributions or as ordinary functions
after smearing in the remaining variables.  The physical value on \(I\) is the
upper-edge boundary value \(\mathcal M_{ab}^{\rm I}(s+\ii0)\).

The adjacent second-sheet germ relative to the \(\phi_1\phi_1\) channel is the
analytic germ obtained by continuing \(\mathcal M_{ab}^{\rm I}\) from the
upper edge of \(I\) through the cut once.  Equivalently, in a local logarithmic
coordinate for the cut, the second-sheet continuation replaces the lower
first-sheet boundary value
\[
  \log(-r+\ii0)=\log r+\ii\pi,\qquad r>0,
\]
by the continued value
\[
  \log_{\rm II}(-r)=\log r-\ii\pi .
\]
For a square-root channel variable this is the same operation as changing the
sign of the phase-space square root \(\rho_1\) relative to the first sheet.
A resonance in this channel is a pole of this adjacent germ, not a pole of the
physical upper-edge boundary value and not an additional normalizable
asymptotic state.
\end{definition}

For the scalar model above the first-sheet exchange amplitude is
\(\mathcal M_s^{\rm I}(s)=g^2/F_{\rm I}(s)\), with \(F_{\rm I}\) given by the
Feynman boundary prescription in the preceding section.  The first sheet
contains the physical branch cut beginning at \(4m_1^2\).  The denominator
\(F_{\rm II}\) denotes the continuation of \(F_{\rm I}\) through this cut in
the sense of
Definition~\ref{def:adjacent-second-sheet-single-channel}.  On a real point
of the cut, the logarithmic discontinuity is
\[
  \log(-|a|-\ii0)-\log(-|a|+\ii0)=-2\pi\ii .
\]
Therefore the lower second-sheet boundary differs from the lower first-sheet
boundary by the finite discontinuity generated by this logarithm.  Near
\(s=M_R^2\), the branch inherited by \(F_{\rm II}\) has
\[
  F_{\rm II}(s)
  =
  M_R^2-s-\ii W(s)+O(g^4,g^2(s-M_R^2)),
\]
where \(W(s)=g^2\rho_1(s)/(32\pi)\).  This is the local analytic reason a pole
can occur below the real axis on the adjacent sheet while no first-sheet pole
appears at the same point.

\begin{proposition}[No first-sheet resonance zero in the scalar model]
\label{prop:no-first-sheet-resonance-zero-scalar-model}
Use the first-sheet branch of
\[
  F_{\rm I}(s)
  =
  -s+(m_2^R)^2
  +
  \frac{g^2}{32\pi^2}
  \int_0^1\dd x\,
  \log\left(1-\frac{s\,x(1-x)}{m_1^2}\right)
\]
defined by continuation from the real interval \(s<4m_1^2\), with the cut
placed along \(s\ge4m_1^2\).  Then \(F_{\rm I}\) has no zero in the upper
half of the first sheet and no zero in the lower half of the first sheet.
\end{proposition}

\begin{proof}
Let \(s=u+\ii v\) with \(v>0\).  For \(0<x<1\), set
\(\alpha=x(1-x)/m_1^2>0\).  The logarithm has argument
\[
  1-\alpha s=(1-\alpha u)-\ii\alpha v,
\]
whose imaginary part is strictly negative.  On the branch reached from
\(s<4m_1^2\), its argument lies in \((-\pi,0)\); hence
\[
  \operatorname{Im}\log(1-\alpha s)<0 .
\]
At the endpoints \(x=0,1\) the logarithm is zero; they do not affect the
integral.  Also
\[
  \operatorname{Im}(-s)=-v<0 .
\]
Since \(g^2/(32\pi^2)>0\), every nonconstant contribution to
\(\operatorname{Im}F_{\rm I}(s)\) has negative sign, and the integral is
strictly negative.  Therefore \(\operatorname{Im}F_{\rm I}(s)<0\) in the
upper half-plane, so \(F_{\rm I}(s)\neq0\) there.  The lower half-plane
statement follows either by the same argument with \(v<0\), giving
\(\operatorname{Im}F_{\rm I}(s)>0\), or by Schwarz reflection of the
first-sheet branch.  Thus the zero associated with the resonance cannot lie
on either half of the first sheet.
\end{proof}

\begin{proposition}[Narrow pole on the adjacent sheet]
\label{prop:second-sheet-narrow-pole-newton}
Let
\[
  W_R=M_R\Gamma_{\rm cut}(M_R)
  =
  \frac{g^2}{32\pi}
  \sqrt{1-\frac{4m_1^2}{M_R^2}}
  +O(g^4),
\]
and assume \(M_R^2>4m_1^2\) is separated from the threshold by a fixed
positive amount.  Suppose that in a fixed neighborhood of \(M_R^2\) the
second-sheet denominator has the weak-coupling expansion
\[
  F_{\rm II}(s)
  =
  M_R^2-s-\ii W_R
  +(s-M_R^2)R_1(s,g)+R_0(g),
  \qquad
  R_1=O(g^2),\quad R_0=O(g^4),
\]
with \(R_1\) holomorphic and uniformly bounded with its first derivative.
Then, for sufficiently small \(g\), there is a unique zero of
\(F_{\rm II}\) in a disk \(|s-M_R^2|\le Cg^2\), and it is
\[
  s_\ast
  =
  M_R^2-\ii M_R\Gamma_{\rm cut}(M_R)+O(g^4).
\]
\end{proposition}

\begin{proof}
Write \(s=M_R^2+\delta\).  The equation \(F_{\rm II}(s)=0\) is
\[
  \delta
  =
  -\ii W_R+\delta R_1(M_R^2+\delta,g)+R_0(g).
\]
For \(|\delta|\le Cg^2\) and \(C\) large enough, the right-hand side maps the
disk to itself for all sufficiently small \(g\), because
\(W_R=O(g^2)\), \(\delta R_1=O(g^4)\), and \(R_0=O(g^4)\).  Its derivative
with respect to \(\delta\) is
\[
  R_1(M_R^2+\delta,g)
  +
  \delta\,\partial_s R_1(M_R^2+\delta,g),
\]
which is \(O(g^2)\) uniformly on the disk.  Thus the map is a contraction,
and it has a unique fixed point.  The fixed-point equation immediately gives
\[
  \delta=-\ii W_R+O(g^4),
\]
which is the stated pole location.  Equivalently,
\(F_{\rm II}(M_R^2)=-\ii W_R+O(g^4)\) and
\(F_{\rm II}'(M_R^2)=-1+O(g^2)\), so the same result is the controlled
one-step Newton approximation.
\end{proof}

Equivalently, the pole width defined by
\(E_\ast=\sqrt{s_\ast}=M_{\rm pole}-\ii\Gamma_{\rm pole}/2\) obeys
\[
  M_{\rm pole}\Gamma_{\rm pole}
  =
  M_R\Gamma_{\rm cut}(M_R)+O(g^4).
\]
More invariantly, \(s_\ast\) is defined by the denominator obtained by
continuing the first-sheet \(F_{\rm I}\) through the elastic cut:
\[
  F_{\mathrm{II}}(s_\ast)=0 .
\]
The number \(s_\ast\) is an analytic datum of the stable-particle scattering
amplitude. Its position sets the resonance location and the pole lifetime
scale in the narrow-width regime; it is not the same kind of object as a
physical-axis line-shape width unless the narrow-resonance hypotheses of the
preceding section hold.

The same distinction is visible in the spectral representation. A stable
one-particle state gives an atom in the spectral measure and supplies a
Haag--Ruelle asymptotic sector. An open channel gives continuous spectral
measure beginning at threshold. A narrow resonance is a pronounced analytic
feature of that continuum: it can appear as a peak in the spectral density and
as a pole after analytic continuation, while the external scattering states
remain the stable states used in the construction of the \(S\)-matrix.

\section{Unstable Resonances and External-Leg Reduction}

The LSZ theorem applies to an external particle only when the theory supplies
an isolated real mass shell in the physical Hilbert space.  In the notation of
Chapter~\ref{ch:lsz-reduction}, this means an orthogonal projection
\[
  P_1=E(\Sigma_m^+)
\]
onto a Wigner one-particle subspace, a positive pole residue in a local-field
two-point function, and Haag--Ruelle incoming and outgoing states built from
that subspace.  A resonance pole
\(s_\ast=M_{\rm pole}^2-\ii M_{\rm pole}\Gamma_{\rm pole}
-\Gamma_{\rm pole}^2/4\) has
none of these properties.  It is a pole of an analytically continued stable
scattering amplitude on a nonphysical sheet; the associated outgoing Gamow
functional is not a normalizable vector in \(\Hilb\), and there is no
projection-valued spectral atom \(E(\{p:p^2=-s_\ast\})\).

Therefore a residue at \(s=s_\ast\) is not an LSZ external-leg residue.
Physical \(S\)-matrix elements with a resonance are matrix elements between
stable asymptotic states.  Near an isolated resonance pole the meromorphic
continuation of such a stable-particle amplitude may have the local form
\[
  \mathcal M_{a\to b}(s,\alpha,\beta)
  =
  \frac{r_{a}(\alpha)\,r_{b}(\beta)}{s_\ast-s}
  +
  \mathcal M_{\rm reg}(s,\alpha,\beta),
\]
where \(a,b\) label stable channels and \(\alpha,\beta\) denote the remaining
kinematic variables.  The functions \(r_a,r_b\) are pole residues of the
continued stable-channel amplitude.  They can be used to describe production
and decay through the resonance, but they do not define an incoming or
outgoing one-resonance Hilbert-space state.

Perturbation theory often packages this pole information in a
Stueckelberg--Veltman, or complex-mass-shell, prescription.  One analytically
continues the inverse dressed propagator to the pole,
\[
  F_{\rm II}(s_\ast)=0,
  \qquad
  Z_\ast^{-1}= -F_{\rm II}'(s_\ast),
\]
and replaces the resonant propagator near the pole by
\[
  \frac{g^2}{F_{\rm II}(s)}
  =
  \frac{g^2 Z_\ast}{s_\ast-s}
  +
  \text{terms holomorphic near }s_\ast .
\]
This is a pole-scheme parametrization of the stable-channel amplitude, not a
new LSZ theorem.  Its usefulness requires a controlled narrow-resonance
regime: the pole must be isolated from other nearby poles and branch points,
the width scale \(M_{\rm pole}\Gamma_{\rm pole}\) must be small compared with
the squared-mass distance to thresholds or singularities whose variation is
being neglected, and the nonresonant background must be slowly varying on the
same scale.  If the pole lies close to a threshold, or if several channels
produce nearby singularities, the simple Breit--Wigner or complex-mass-shell
description is replaced by the corresponding multi-channel analytic line
shape.  The external states remain the stable states in all cases.

\section{Pole-Factorized Resonance Objects}

The preceding discussion leaves a precise mathematical problem.  Stable
external particles are vectors in asymptotic Hilbert spaces.  Resonances are
singularities of analytic continuations of matrix elements between such
vectors.  A candidate ``amplitude with an external resonance'' must therefore
be extracted from stable-channel amplitudes by pole residues.

\begin{definition}[Local pole neighborhood]
Fix stable channels \(a,b\) and kinematic variables
\((s,\alpha,\beta)\), where \(s\) is the invariant mass squared of a channel
whose analytic continuation has a pole.  A local pole neighborhood consists
of a sheet \(\mathfrak s\) of the analytically continued amplitude
\(\mathcal M^{\mathfrak s}_{a\to b}(s,\alpha,\beta)\), an open set
\(U\subset\C\) containing \(s_\ast\), and open real or complex neighborhoods
\(V_a,V_b\) of \(\alpha,\beta\), such that the only singularity in
\((U\setminus\{s_\ast\})\times V_a\times V_b\) is a simple pole at
\(s=s_\ast\).
\end{definition}

\begin{definition}[Pole-factorized resonance object]
\label{def:pole-factorized-resonance-object}
In a local pole neighborhood, the pole-residue kernel is
\begin{equation}
  R_{ab}(\alpha,\beta)
  =
  \lim_{s\to s_\ast}
  (s_\ast-s)\,
  \mathcal M^{\mathfrak s}_{a\to b}(s,\alpha,\beta).
  \label{eq:resonance-residue-kernel}
\end{equation}
If the residue kernel has rank one in the channel variables, a
pole-factorized resonance object is an equivalence class of factorizations
\[
  R_{ab}(\alpha,\beta)=r_{b}(\beta)\,r_{a}(\alpha),
  \qquad
  (r_a,r_b)\sim(\lambda r_a,\lambda^{-1}r_b),
  \qquad
  \lambda\in\C^\times .
\]
The functions \(r_a,r_b\) are production and decay residue functionals.  The
equivalence class, rather than either factor separately, is the invariant
object in the one-pole problem.
\end{definition}

For a resonance in a spin-\(j\) channel, the residue kernel carries the
indices of the corresponding partial wave or covariant tensor structure.
After projecting to the irreducible spin component at the pole, rank one means
rank one in the remaining channel space.  In gauge theory this projection must
be performed on the gauge-invariant stable-channel amplitude, or
equivalently on BRST cohomology representatives whose pole residue is
independent of the chosen gauge-fixing representatives.  Pole position and
full stable-channel residues are the primary objects; a split into
``production'' and ``decay'' factors is meaningful only after the
normalization equivalence in
Definition~\ref{def:pole-factorized-resonance-object} has been imposed.

The same construction extends to several resonance variables only under a
strong isolation hypothesis.  Suppose a stable-channel amplitude depends on
complex variables \(s_1,\ldots,s_N\), and the chosen sheet contains simple
poles at \(s_i=s_{\ast i}\).  If there is a polydisc
\(\prod_i U_i\) in which the only singular divisors are these \(N\) simple
pole hyperplanes, define the nested residue
\begin{equation}
  R^{(N)}
  =
  \lim_{s_N\to s_{\ast N}}\cdots
  \lim_{s_1\to s_{\ast1}}
  \left(\prod_{i=1}^N(s_{\ast i}-s_i)\right)
  \mathcal M^{\mathfrak s}(s_1,\ldots,s_N,\gamma).
  \label{eq:nested-resonance-residue}
\end{equation}
The order of residues is then immaterial by meromorphy in the polydisc.  This
object is the mathematical candidate for an amplitude with \(N\) external
resonance slots and any remaining stable external data \(\gamma\).  The
definition collapses when a threshold branch point, an anomalous threshold, or
another pole intersects the same neighborhood.

\begin{proposition}[Properties inherited from stable-channel scattering]
\label{prop:resonance-residue-inherited-properties}
Assume the stable-channel \(S\)-matrix is unitary on the physical boundary,
Poincare covariant, and obtained from local fields satisfying the analyticity
hypotheses used in the preceding chapters.  Let
\(R_{ab}\) be the pole-residue kernel in a local pole neighborhood.  Then:
\begin{enumerate}
\item \(s_\ast\) and \(R_{ab}\) are invariant under changes of interpolating
fields that preserve the stable-channel \(S\)-matrix.
\item covariance of the stable amplitude induces covariance of \(R_{ab}\) in
the little-group or partial-wave representation labeling the pole.
\item physical unitarity gives reality and conjugation constraints relating
the pole on the chosen sheet to the conjugate singularity on the reflected
sheet, and it fixes the positive inclusive decay rate in the narrow isolated
regime through the discontinuity of the stable-channel amplitude.
\item in a gauge theory, if the stable-channel amplitude is represented by
BRST-closed external operators, the residue kernel is independent of BRST-exact
changes of those representatives.
\end{enumerate}
\end{proposition}

\begin{proof}
The residue kernel is obtained by applying the linear functional
\(\operatorname{Res}_{s=s_\ast}\) to a meromorphic continuation of a
stable-channel matrix element.  Field redefinitions that leave the
stable-channel \(S\)-matrix unchanged leave this meromorphic germ unchanged,
and hence leave its residue unchanged.  Poincare transformations act on the
stable amplitude before continuation; since analytic continuation and taking a
simple residue commute with the finite-dimensional tensor or partial-wave
projection, the residue transforms in the corresponding representation.

On the physical boundary, unitarity gives \(S^\dagger S=1\), or equivalently
the discontinuity relation for \(T=S-1\) derived from stable-channel
unitarity.  Continuing this relation
through the cut relates the pole on one sheet to the conjugate analytic
structure on the reflected sheet.  In the narrow isolated regime, projecting
the discontinuity onto the pole term gives the same inclusive rate as the
self-energy cut calculation of the previous section.  Finally, a BRST-exact
change of a stable external representative changes a physical stable-channel
matrix element by zero; applying the residue functional to zero gives zero.
\end{proof}

\section{Gamow Functionals and the Boundary of the Construction}

A rigged-Hilbert formulation packages the pole residue into generalized
vectors.  Choose dense test spaces
\[
  \Phi_-\subset\Hilb\subset\Phi_-^\times,
  \qquad
  \Phi_+\subset\Hilb\subset\Phi_+^\times
\]
whose energy wavefunctions have the analyticity needed to continue the
incoming and outgoing scattering amplitudes to the pole.  A decaying Gamow
functional associated with \(s_\ast\) is an element of one dual space whose
pairing with test vectors reproduces the pole residue.  In the rest frame,
with \(E_\ast=M_{\rm pole}-\ii\Gamma_{\rm pole}/2\), its time evolution is a
semigroup:
\[
  \ee^{-\ii Ht}\ket{G_\ast}
  =
  \ee^{-\ii M_{\rm pole}t}\ee^{-\Gamma_{\rm pole}t/2}\ket{G_\ast},
  \qquad t\ge0,
\]
as a statement about pairings with the chosen test space.  The restriction
\(t\ge0\) is part of the analytic test-space construction; a self-adjoint
Hamiltonian on \(\Hilb\) still generates a unitary group on normalizable
vectors.

This rigged-Hilbert language is useful because it records the decaying pole
term without pretending that the resonance is a vector in \(\Hilb\).  It is
also additional structure.  Different choices of test spaces can represent
the same pole residue, and a general local QFT does not come equipped with a
canonical Hardy-space rigging for every multiparticle channel.  For this
monograph the invariant data remain the pole position, sheet, and residue
kernels of stable-channel amplitudes; Gamow functionals are a representation
of those data when the required analytic rigging has been supplied.

\section{Obstructions to External-Resonance Amplitudes}

The pole-residue construction has sharp failure modes.

\begin{enumerate}
\item \textbf{Threshold collision.}  If \(s_\ast\) lies at a distance
comparable to its width from a branch point, there is no neighborhood in which
the amplitude is a single simple pole plus a holomorphic remainder.  The
observable line shape is controlled by the pole and the branch point together.
\item \textbf{Coupled channels.}  With several open channels, the analytic
continuation lives on a multi-sheeted surface.  A pole may be close on one
sheet and absent or shadowed on another.  The residue is then a matrix in
channel space, and rank-one factorization can fail.
\item \textbf{Overlapping resonances.}  If two poles lie in the same local
domain, the residue decomposition depends on the chosen basis in the
finite-dimensional pole subspace unless the pole projectors are separated by
contours.
\item \textbf{Gauge and line-shape conventions.}  A complex pole of the full
stable-channel amplitude is invariant, but masses, widths, branching
fractions, and production factors extracted from a fitted real-axis line shape
depend on the convention used to separate pole and background terms.
\item \textbf{Long-range and massless effects.}  If the stable channels have
infraparticle behavior or uncancelled soft radiation, the stable-channel
amplitude itself must be replaced by an inclusive or dressed observable before
any resonance residue is defined.
\end{enumerate}

\begin{openproblem}[External resonances]
The local construction above gives a candidate object in the isolated
simple-pole regime.  The open problem is to promote these local residues to a
general theory of amplitudes with resonance slots.  Such a theory must specify
the class of local QFTs and channels for which the required analytic
continuations exist, define the allowed rigged-Hilbert test spaces or avoid
them in favor of purely meromorphic residue data, prove compatibility with
stable-channel unitarity, formulate covariance for spin and gauge theories,
and give precise replacement rules when thresholds, overlapping poles,
massless sectors, or coupled-channel sheets obstruct rank-one pole
factorization.  Perturbative complex-mass and pole schemes are calculational
realizations of the isolated-pole part of this problem; they are not a
nonperturbative external-state construction.
\end{openproblem}
