\chapter{The Schwinger Model: Two-Dimensional Quantum Electrodynamics}
\label{ch:schwinger-model}
\ConventionsLorentzian

\section{Definition, Gauge Group, and Observable Sector}

The Schwinger model is quantum electrodynamics in \(1+1\) spacetime
dimensions with one massless Dirac field.  We use coordinates
\(x^\mu=(t,x)\),
\[
  \eta_{\mu\nu}=\operatorname{diag}(-,+),
  \qquad
  \epsilon^{01}=+1,
  \qquad
  \gamma=\gamma^0\gamma^1 .
\]
The two-dimensional spinor conventions are those of
Section~\ref{sec:spinor-conventions}.  Let \(a_\mu\) be a real Abelian
gauge representative and let \(e>0\) be the charge of the dynamical Dirac
field.  The compact \(U(1)\) connection is \(e\,a_\mu\); this convention is
important because external probes will be allowed to have charge \(Q\) with
respect to \(a_\mu\).  Thus
\[
  D_\mu\psi=\partial_\mu\psi-\ii e a_\mu\psi,
  \qquad
  F_{\mu\nu}=\partial_\mu a_\nu-\partial_\nu a_\mu,
  \qquad
  E=F_{01}.
\]
The exact bosonized scalar below coordinates current and line sectors; it is
not a local order parameter for a continuous symmetry.  This interpretation
matches the Coleman boundary in
Section~\ref{sec:low-dimensional-continuous-ssb-no-go}: the gauge-invariant
derivative current algebra can be local and exact while an unsmeared massless
scalar expectation is not a local order parameter of the \(1+1\)-dimensional
QFT.
The local representative Lagrangian is
\begin{equation}
  \mathcal L_{\rm Sch}
  =
  -\frac14F_{\mu\nu}F^{\mu\nu}
  -\bar\psi\gamma^\mu D_\mu\psi
  =
  \frac12E^2-\bar\psi\gamma^\mu\partial_\mu\psi
  +e\,a_\mu J^\mu ,
  \label{eq:schwinger-lagrangian}
\end{equation}
where
\begin{equation}
  J^\mu=\ii\bar\psi\gamma^\mu\psi,
  \qquad
  J_A^\mu=\ii\bar\psi\gamma^\mu\gamma\psi .
  \label{eq:schwinger-vector-axial-currents}
\end{equation}
With these definitions \(eJ^\mu\) is the dynamical electric current.  Gauge
representatives are related by
\[
  a_\mu\mapsto a_\mu+\partial_\mu\xi,
  \qquad
  \psi\mapsto \ee^{\ii e\xi}\psi,
  \qquad
  \bar\psi\mapsto\bar\psi\,\ee^{-\ii e\xi}.
\]
On a spatial circle of circumference \(L\), the gauge group includes large
transformations with \(\xi(x+L)-\xi(x)=2\pi n/e\).  Equivalently, the
holonomy \(\exp(\ii e\oint a_x\dd x)\) is the compact gauge-invariant
zero-mode coordinate.  The local computations below may be performed on the
line and then lifted to the circle by including the compact zero mode and
the large-gauge identifications.

A possible theta density is
\begin{equation}
  \mathcal L_\theta=\frac{e\theta}{2\pi}E .
  \label{eq:schwinger-theta-density}
\end{equation}
On a closed Euclidean two-manifold the curvature of \(e\,a\) has integral
periods, so \(\theta\sim\theta+2\pi\).  In the massless theory an anomalous
axial rotation shifts \(\theta\), and the local spectrum is independent of
\(\theta\).  In the massive deformation this shift is no longer a symmetry,
and \(\theta\) becomes a physical parameter.

\begin{definition}[Probe-charge convention]
\label{def:schwinger-probe-charge-convention}
  A static external probe of charge \(Q\) is defined by the conserved
  external current \(j^\mu_{\rm ext}\) coupled through
  \[
    S_{\rm probe}=\int a_\mu j^\mu_{\rm ext}\,\dd^2x .
  \]
  A straight worldline carries the phase
  \(\exp(\ii Q\int a_\mu\dd x^\mu)\).  The dynamical Dirac field has
  \(Q=e\).  A probe with \(Q/e\in\mathbb Z\) can be screened by dynamical
  charged particles; a probe with nonintegral \(Q/e\) is a fractional probe
  relative to the dynamical charge lattice.
\end{definition}

The canonical momentum conjugate to \(a_x\) is \(E\), while \(a_0\) is a
Lagrange multiplier.  Including an external charge density
\(\rho_{\rm ext}=j^0_{\rm ext}\), variation of \(a_0\) gives the operator
Gauss law
\begin{equation}
  \partial_xE+eJ^0+\rho_{\rm ext}=0 .
  \label{eq:schwinger-gauss-law}
\end{equation}
This equation is a constraint on states and on correlation functions with
line insertions.  It is not optional data added after quantization; it is the
local gauge-invariance condition of the theory.

\begin{definition}[Local sector]
\label{def:schwinger-local-sector}
  The local gauge-invariant sector of the massless Schwinger model is the
  local operator algebra generated, after renormalization, by \(F_{\mu\nu}\),
  by gauge-invariant currents such as \(J^\mu\), and by neutral composite
  fields.  The bare fermion field \(\psi\) is not a gauge-invariant local
  observable.  Charged sectors are described by line-dressed charged
  insertions or by states on a spatial manifold with prescribed electric
  flux at infinity.  The exact solution below is first a solution of the
  local gauge-invariant sector and of specified external line probes.
\end{definition}

\section{Current Algebra, Anomaly, and the Electric-Field Equation}

In two dimensions the vector and axial currents are Hodge dual.  With the
gamma-matrix and orientation conventions fixed above,
\begin{equation}
  J_A^\mu=-\epsilon^{\mu\nu}J_\nu .
  \label{eq:schwinger-current-duality}
\end{equation}
The vector current is exactly conserved as a consequence of gauge symmetry:
\[
  \partial_\mu J^\mu=0 .
\]
The axial current is not conserved.  The two-dimensional anomaly formula of
Chapter~\ref{ch:volume-ii-20}, applied to the background connection
\(A_\mu=e a_\mu\), gives the renormalized operator identity
\begin{equation}
  \partial_\mu J_A^\mu
  =
  \frac{e}{2\pi}\epsilon^{\mu\nu}F_{\mu\nu}
  =
  \frac{e}{\pi}E .
  \label{eq:schwinger-axial-anomaly}
\end{equation}
The identity includes the finite contact term chosen in the definition of
the axial-vector current correlator.  It is not the Euler-Lagrange equation
of the classical massless Dirac action.

\begin{proposition}[Anomaly-induced massive electric field]
\label{prop:schwinger-anomaly-induced-electric-mass}
  In the absence of external line insertions, the gauge-invariant electric
  field of the massless Schwinger model obeys
  \begin{equation}
    (\Box-m_{\rm Sch}^2)E=0,
    \qquad
    m_{\rm Sch}^2=\frac{e^2}{\pi},
    \qquad
    \Box=\partial_\mu\partial^\mu .
    \label{eq:schwinger-electric-field-mass-equation}
  \end{equation}
\end{proposition}

\begin{proof}
  The Maxwell equation obtained from \eqref{eq:schwinger-lagrangian} is the
  renormalized local operator identity
  \[
    \partial_\nu F^{\nu\mu}+eJ^\mu=0 .
  \]
  Here it is tested in the sector with no external line insertion.  With
  external Wilson lines, the right side would acquire the corresponding
  distributional endpoint or worldline current, and the homogeneous equation
  below would be replaced by an inhomogeneous screened-field equation.
  Since \(F^{10}=E\) and \(F^{01}=-E\), its two components are
  \[
    \partial_xE+eJ^0=0,
    \qquad
    -\partial_tE+eJ^1=0 .
  \]
  Hence
  \[
    J^0=-\frac1e\partial_xE,
    \qquad
    J^1=\frac1e\partial_tE .
  \]
  Using \eqref{eq:schwinger-current-duality},
  \(J_A^0=-J^1\) and \(J_A^1=-J^0\).  Therefore
  \[
    \partial_\mu J_A^\mu
    =
    -\partial_tJ^1-\partial_xJ^0
    =
    \frac1e(-\partial_t^2+\partial_x^2)E
    =
    \frac1e\Box E .
  \]
  Comparing this distributional consequence of Maxwell's equation with the
  anomaly identity \eqref{eq:schwinger-axial-anomaly}, with the same local
  contact-term convention for the axial current, gives
  \(\Box E/e=eE/\pi\), which is \eqref{eq:schwinger-electric-field-mass-equation}.
\end{proof}

This derivation uses only gauge invariance, the exact current anomaly, and
the Maxwell equation.  It already proves that the local electric field
creates a massive neutral particle.  Bosonization gives a stronger statement:
it identifies the entire current-electric-field sector as a free massive
scalar theory and gives exact formulas for line-probe response.

\section{Bosonization of the Current Sector}

The conservation law for \(J^\mu\) implies locally that \(J^\mu\) is the
Hodge dual of a derivative.  The normalization is not arbitrary; it is fixed
by the current algebra of one Dirac fermion.

\begin{theorem}[Current-sector bosonization]
\label{thm:schwinger-current-sector-bosonization}
  Let the free massless Dirac field be quantized in the vacuum sector with
  the current normalization \eqref{eq:schwinger-vector-axial-currents}.  The
  gauge-invariant Abelian current subalgebra is represented by the derivative
  algebra of a real scalar field \(\varphi\), normalized by
  \[
    \mathcal L_\varphi
    =
    -\frac12\partial_\mu\varphi\,\partial^\mu\varphi,
    \qquad
    [\widehat\varphi(t,x),\partial_0\widehat\varphi(t,y)]
    =
    \ii\delta(x-y),
  \]
  through
  \begin{equation}
    J^\mu=\frac1{\sqrt\pi}\epsilon^{\mu\nu}\partial_\nu\varphi,
    \qquad
    J_A^\mu=-\frac1{\sqrt\pi}\partial^\mu\varphi .
    \label{eq:schwinger-bosonization-currents}
  \end{equation}
  Equivalently, after the same local contact-term convention is fixed on
  both sides, the current-generating functional satisfies
  \begin{equation}
    Z_{\rm Dirac}[b]
    =
    \int\mathcal D\varphi\,
    \exp\!\left\{
      \ii\int\dd^2x
      \left[
        -\frac12\partial_\mu\varphi\partial^\mu\varphi
        +\frac1{\sqrt\pi}b_\mu\epsilon^{\mu\nu}\partial_\nu\varphi
      \right]
    \right\}.
    \label{eq:schwinger-current-generating-functional}
  \end{equation}
  This is the Lorentzian continuation of the Euclidean current-generating
  functional.  When the local
  current algebra is extended by charge-one fermion vertex operators, the
  scalar is compact with period \(\sqrt\pi\); this compactification data is
  invisible to the derivative current algebra alone.
\end{theorem}

\begin{proof}
The first step fixes the normalization independently of any path-integral
argument.  The equal-time commutator of the Dirac current, defined by
gauge-invariant point splitting and with the vector Ward identity preserved,
is
  \begin{equation}
    [\widehat J^0(t,x),\widehat J^1(t,y)]
    =
    -\frac{\ii}{\pi}\partial_x\delta(x-y).
    \label{eq:schwinger-current-schwinger-term}
  \end{equation}
  This Schwinger term follows directly from the short-distance singularity
  of the two chiral components of a free massless Dirac field.  If
  \eqref{eq:schwinger-bosonization-currents} holds, then
  \[
    J^0=\frac1{\sqrt\pi}\partial_x\varphi,
    \qquad
    J^1=-\frac1{\sqrt\pi}\partial_t\varphi ,
  \]
  and the canonical scalar commutator gives exactly
  \eqref{eq:schwinger-current-schwinger-term}.  Hence the coefficient in
  \eqref{eq:schwinger-bosonization-currents} is forced.

It remains to prove that the same scalar gives all current correlators.  Work
first in Euclidean signature on the plane, and couple a real Schwartz one-form
\(b=b_\mu\dd x^\mu\) to the conserved vector current.  The Hodge decomposition
on \(\R^2\) gives
\[
  b_\mu=\partial_\mu\lambda+\epsilon_{\mu\nu}\partial^\nu\rho,
\]
up to harmonic zero modes, which are absent for Schwartz sources.  The vector
Ward identity removes the exact part \(\partial_\mu\lambda\), so the
generating functional depends on \(\rho\) and on local contact conventions
only.  The transverse coupling can be written as an axial pure gauge:
\[
  b_\mu J^\mu
  =
  \epsilon_{\mu\nu}\partial^\nu\rho\,J^\mu
  =
  -\partial_\mu\rho\,J_A^\mu,
\]
with the current convention
\eqref{eq:schwinger-vector-axial-currents}.  A local axial rotation
\(\psi\mapsto e^{\ii\rho\gamma_5}\psi\) removes this coupling from the free
Dirac action.  The entire change of the determinant is the Fujikawa Jacobian
for the same contact-term convention that gives
\eqref{eq:schwinger-axial-anomaly}; in the present normalization,
\[
  \log\frac{Z_{\rm Dirac}[b]}{Z_{\rm Dirac}[0]}
  =
  -\frac1{2\pi}\int_{\R^2}\dd^2x\,\partial_\mu\rho\,\partial^\mu\rho .
\]
Equivalently, in momentum space,
  \begin{equation}
    W_E[b]
    =
    \frac1{2\pi}
    \int\frac{\dd^2p}{(2\pi)^2}\,
    b_\mu(-p)
    \left(\delta^{\mu\nu}-\frac{p^\mu p^\nu}{p^2}\right)
    b_\nu(p).
    \label{eq:schwinger-euclidean-current-functional}
  \end{equation}
where \(Z_{\rm Dirac}[b]=\exp[-W_E[b]]Z_{\rm Dirac}[0]\).  This derivation is
exact: after the source is removed by the axial rotation, there is no
remaining interaction, and the anomaly Jacobian is quadratic in \(\rho\).

The Euclidean scalar representative gives the same functional.  With action
\[
  S_E[\varphi;b]
  =
  \int\dd^2x
  \left[
    \frac12\partial_\mu\varphi\partial^\mu\varphi
    -\frac{\ii}{\sqrt\pi}b_\mu\epsilon^{\mu\nu}\partial_\nu\varphi
  \right],
\]
the Gaussian integral over \(\varphi\), modulo the constant zero mode, gives
\eqref{eq:schwinger-euclidean-current-functional}, because
  \[
    \epsilon^{\mu\alpha}p_\alpha
    \epsilon^{\nu\beta}p_\beta
    =
    p^2\delta^{\mu\nu}-p^\mu p^\nu .
  \]
Functional differentiation of the identical Gaussian source functional gives
all Euclidean current distributions, with the already fixed local contact
terms.  Analytic continuation gives
\eqref{eq:schwinger-current-generating-functional} and the Lorentzian current
formula \eqref{eq:schwinger-bosonization-currents}.

The compact period is fixed only after adjoining charged fields.  A charge-one
fermion operator is represented by a vertex operator whose monodromy with the
current has unit charge; with the current normalization above, the fermion
mass operator has the local representative
\(\cos(2\sqrt\pi\,\varphi)\).  Single-valuedness of this local neutral
operator gives \(\varphi\sim\varphi+\sqrt\pi\).  This compactification datum
does not enter the local derivative current correlators, which are the only
bosonization data used in the gauge-invariant Schwinger-model calculations
below.
\end{proof}

\section{Exact Bosonized Gauge Theory}

\begin{lemma}[Exact bosonized local gauge sector]
\label{lem:schwinger-exact-bosonized-gauge-sector}
In the massless Schwinger model, the gauge-invariant local current and
electric-field sector is represented by the bosonic Lagrangian
\begin{equation}
  \mathcal L_{\rm bos}
  =
  -\frac12\partial_\mu\varphi\,\partial^\mu\varphi
  -\frac14F_{\mu\nu}F^{\mu\nu}
  +\frac{e}{\sqrt\pi}a_\mu\epsilon^{\mu\nu}\partial_\nu\varphi
  +\frac{e\theta}{2\pi}E .
  \label{eq:schwinger-bosonized-action}
\end{equation}
On a fixed local electric-flux branch this representative is equivalent to
\begin{equation}
  \mathcal L_{\rm eff}
  =
  -\frac12\partial_\mu\varphi\,\partial^\mu\varphi
  -\frac12\,\frac{e^2}{\pi}
  \left(\varphi+\frac{\theta}{2\sqrt\pi}\right)^2 .
  \label{eq:schwinger-effective-massive-scalar}
\end{equation}
The electric field is the massive scalar field
\begin{equation}
  E
  =
  -\frac{e}{\sqrt\pi}
  \left(\varphi+\frac{\theta}{2\sqrt\pi}\right),
  \qquad
  (\Box-m_{\rm Sch}^2)E=0,
  \qquad
  m_{\rm Sch}^2=\frac{e^2}{\pi}.
  \label{eq:schwinger-electric-field-scalar-relation}
\end{equation}
\end{lemma}

\begin{proof}
Set \(b_\mu=e a_\mu\) in the current-sector identity
\eqref{eq:schwinger-current-generating-functional}.  Adding the Maxwell term
and the theta density \eqref{eq:schwinger-theta-density} gives
\eqref{eq:schwinger-bosonized-action}.  Since
\(-\frac14F_{\mu\nu}F^{\mu\nu}=\frac12E^2\) in the mostly-plus convention and
\[
  a_\mu\epsilon^{\mu\nu}\partial_\nu\varphi
  =
  a_0\partial_x\varphi-a_x\partial_t\varphi
  =
  E\varphi+\partial_\mu(\cdots),
\]
the same local branch is represented, up to a boundary term, by
\begin{equation}
  \mathcal L_{\rm bos}
  =
  -\frac12\partial_\mu\varphi\,\partial^\mu\varphi
  +\frac12E^2
  +E\left(\frac{e}{\sqrt\pi}\varphi+\frac{e\theta}{2\pi}\right).
  \label{eq:schwinger-electric-field-form}
\end{equation}
In two spacetime dimensions \(E\) carries no independent transverse
propagating gauge degree of freedom.  Its local stationary equation is
\[
  E
  =
  -\frac{e}{\sqrt\pi}\varphi-\frac{e\theta}{2\pi}.
\]
Substitution into \eqref{eq:schwinger-electric-field-form} gives
\eqref{eq:schwinger-effective-massive-scalar}.  The Euler-Lagrange equation
from \eqref{eq:schwinger-effective-massive-scalar} is
\[
  \Box\!\left(\varphi+\frac{\theta}{2\sqrt\pi}\right)
  =
  \frac{e^2}{\pi}
  \left(\varphi+\frac{\theta}{2\sqrt\pi}\right),
\]
and multiplying by \(-e/\sqrt\pi\) gives the electric-field equation in
\eqref{eq:schwinger-electric-field-scalar-relation}.
\end{proof}

The lemma is an exact equality for the local current and electric-field
sector of the massless theory.  Gauge-variant fermion fields, compact-boson
vertex operators, and gauge lines belong to the charged-sector data; the local
neutral algebra is the sector described above.

\begin{remark}[Theta periodicity and branches]
  The local expression \eqref{eq:schwinger-effective-massive-scalar} is
  written on one electric-flux branch.  In the compact theory, large gauge
  transformations and the compactness
  \(\varphi\sim\varphi+\sqrt\pi\) identify the branch obtained by
  \(\theta\mapsto\theta+2\pi\) with the branch obtained by shifting
  \(\varphi\).  A formulation on a spatial circle keeps the holonomy zero
  mode and sums over the corresponding integer electric-flux sectors.  The
  massless theory is independent of \(\theta\), because an axial rotation
  shifts \(\theta\) and no fermion mass term obstructs this shift.
\end{remark}

Figure~\ref{fig:schwinger-model-exact-solution} displays two gauge-invariant
consequences of the same calculation: the local electric field has a massive
pole with \(m_{\rm Sch}^2=e^2/\pi\), and the static kernel changes from the
unscreened one-dimensional Coulomb kernel to a screened Yukawa kernel.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \draw[->] (-0.1,0) -- (4.95,0) node[below] {\(|k|/m_{\rm Sch}\)};
    \draw[->] (0,-0.1) -- (0,3.05) node[left] {\(\omega/m_{\rm Sch}\)};
    \draw[gray!70,dashed] plot[domain=0:3.8,samples=2]
      ({\x},{0.68*\x});
    \draw[blue!70!black,thick] plot[domain=0:3.8,samples=80]
      ({\x},{0.68*sqrt(1+\x*\x)});
    \node[blue!70!black,anchor=west] at (2.05,2.15)
      {\(\omega^2=k^2+m_{\rm Sch}^2\)};
    \node[gray!60!black,anchor=west] at (2.45,1.25)
      {\(\omega=|k|\)};
    \draw[blue!70!black,densely dotted] (0,0.68) -- (3.8,0.68);
    \node[blue!70!black,anchor=east] at (-0.05,0.68) {\(1\)};
    \node[align=center] at (2.35,-0.68)
      {local electric-field pole};

    \begin{scope}[shift={(6.25,0)},x=0.86cm,y=0.62cm]
      \draw[->] (-0.1,0) -- (5.55,0) node[below] {\(m_{\rm Sch}R\)};
      \draw[->] (0,-0.1) -- (0,5.25)
        node[left] {\(\frac{2m_{\rm Sch}V}{Q^2}\)};
      \draw[gray!70,dashed] plot[domain=0:5,samples=2]
        ({\x},{\x});
      \draw[red!70!black,thick] plot[domain=0:5,samples=80]
        ({\x},{1-exp(-\x)});
      \draw[red!70!black,densely dotted] (0,1) -- (5,1);
      \draw[densely dotted] (1,0) -- (1,1);
      \node[red!70!black,anchor=west] at (2.15,1.35)
        {\(1-\ee^{-m_{\rm Sch}R}\)};
      \node[gray!60!black,anchor=west] at (2.65,3.05)
        {pure Maxwell: \(m_{\rm Sch}R\)};
      \node[anchor=north] at (1,-0.10) {\(1\)};
      \node[red!70!black,anchor=east] at (-0.08,1) {\(1\)};
      \node[align=center] at (2.75,-1.08)
        {static screening length \(m_{\rm Sch}^{-1}\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{Two observable faces of the same massless Schwinger-model
  calculation.  The anomaly-fixed current algebra gives the electric field a
  massive neutral pole, and the same mass changes the static probe potential
  from a linearly growing Maxwell kernel to the screened form derived below.}
  \label{fig:schwinger-model-exact-solution}
\end{figure}

\section{Static Probes and Screening}

The response to external static charges is a line-observable question, not a
statement about the local particle spectrum alone.  For a neutral pair
\[
  \rho_{\rm ext}(x)
  =
  Q\delta(x-R/2)-Q\delta(x+R/2),
\]
pure Maxwell theory in one spatial dimension gives a constant electric field
between the probes.  With the Gauss law \(\partial_xE+\rho_{\rm ext}=0\) in
the interval between dynamical charges,
\[
  V_{\rm pure}(R)=\frac{Q^2}{2}R .
\]

\paragraph{Screened static potential in the massless model.}
For the neutral static probe pair
\[
  \rho_R(x)=Q\delta(x-R/2)-Q\delta(x+R/2),
  \qquad R\ge0,
\]
the massless Schwinger model gives the static energy normalized by
\(V_{\rm Sch}(0)=0\)
\begin{equation}
  V_{\rm Sch}(R)
  =
  Q^2\left(G_{m_{\rm Sch}}(0)-G_{m_{\rm Sch}}(R)\right)
  =
  \frac{Q^2}{2m_{\rm Sch}}
  \left(1-\ee^{-m_{\rm Sch}R}\right),
  \label{eq:schwinger-screened-static-potential}
\end{equation}
where
\[
  \left(-\partial_x^2+m^2\right)G_m(x)=\delta(x),
  \qquad
  G_m(x)=\frac{\ee^{-m|x|}}{2m}.
\]
Consequently the static force decays exponentially at distance
\(m_{\rm Sch}^{-1}\), and there is no asymptotic linear electric string in
the massless model.

Integrating out the massless current sector changes the static Coulomb kernel
from \((-\partial_x^2)^{-1}\) to
\((-\partial_x^2+m_{\rm Sch}^2)^{-1}\).  Equivalently, the temporal
representative in Coulomb gauge obeys the screened static equation
\[
  \left(-\partial_x^2+m_{\rm Sch}^2\right)a_0(x)=\rho_R(x).
\]
The work needed to assemble the neutral pair from the coincident
configuration is the positive quadratic form
\begin{equation}
  V_{\rm Sch}(R)
  =
  \frac12\int\frac{\dd k}{2\pi}\,
  \frac{|\rho_R(k)|^2}{k^2+m_{\rm Sch}^2},
  \qquad
  \rho_R(k)=Q\bigl(\ee^{-\ii kR/2}-\ee^{\ii kR/2}\bigr).
  \label{eq:schwinger-static-quadratic-kernel}
\end{equation}
Since
\[
  |\rho_R(k)|^2=2Q^2(1-\cos kR),
\]
equation \eqref{eq:schwinger-static-quadratic-kernel} becomes
\[
  Q^2\int\frac{\dd k}{2\pi}
  \frac{1-\cos kR}{k^2+m_{\rm Sch}^2}
  =
  Q^2\left(G_{m_{\rm Sch}}(0)-G_{m_{\rm Sch}}(R)\right).
\]
The displayed Green function follows by solving the ordinary differential
equation away from \(x=0\), imposing decay at infinity, and matching the jump
\(-G_m'(0^+)+G_m'(0^-)=1\).  This gives
\eqref{eq:schwinger-screened-static-potential}.

This screening statement is compatible with
\eqref{eq:schwinger-electric-field-mass-equation}.  The local electric field
creates a massive neutral particle, while external flux is screened because
the induced charge density can rearrange over the length scale
\(m_{\rm Sch}^{-1}\).  These are two different gauge-invariant observables
whose common origin is the same anomaly-fixed current algebra.

\section{Massive Fermion Variant and Fractional-Probe Confinement}

Now add a Dirac mass \(M\bar\psi\psi\).  The axial rotation that removed
\(\theta\) in the massless theory also rotates the mass phase, so the
physical parameter is the relative phase between the mass and the theta
term.  Choosing \(M>0\) real, bosonization gives the sine-Gordon perturbation
\begin{equation}
  \Delta\mathcal L_M
  =
  \kappa M\cos(2\sqrt\pi\,\varphi+\theta).
  \label{eq:massive-schwinger-sine-gordon-perturbation}
\end{equation}
The constant \(\kappa>0\) has mass dimension one and is fixed by the
normalization of the renormalized composite mass operator.  Equivalently,
\(\kappa\) is the magnitude of the chiral condensate in the massless theory
in the corresponding normal-ordering scheme.  The scheme dependence of
\(\kappa\) is exactly compensated by the scheme dependence of the
renormalized mass \(M\).

After eliminating \(E\) on a local branch, the scalar potential is
\begin{equation}
  V_M(\varphi)
  =
  \frac12\,\frac{e^2}{\pi}
  \left(\varphi+\frac{\theta}{2\sqrt\pi}\right)^2
  -\kappa M\cos(2\sqrt\pi\,\varphi+\theta).
  \label{eq:massive-schwinger-effective-potential}
\end{equation}
The quadratic term should again be understood together with the integer flux
branches of the compact gauge theory.  The cosine is periodic under
\(\varphi\mapsto\varphi+\sqrt\pi\); the electric term records which flux
branch is being used.

\begin{proposition}[Fractional-probe string tension at small fermion mass]
\label{prop:massive-schwinger-fractional-string-tension}
Fix the normalization of the renormalized mass operator by
\eqref{eq:massive-schwinger-sine-gordon-perturbation} and take \(M\) small.
A static probe pair with charges \(Q\) and \(-Q\) shifts the effective theta
angle in the interval between the probes by
\begin{equation}
  \theta_{\rm inside}
  =
  \theta+\frac{2\pi Q}{e}
  \quad
  \operatorname{mod} 2\pi .
  \label{eq:schwinger-theta-shift-probe}
\end{equation}
The leading large-\(R\) energy-density shift in that interval is
\begin{equation}
  \Delta\varepsilon_Q(\theta)
  =
  \kappa M
  \left[
    \cos\theta-\cos\!\left(\theta+\frac{2\pi Q}{e}\right)
  \right]
  +O(M^2),
  \label{eq:massive-schwinger-string-tension-theta}
\end{equation}
with the physical branch chosen by minimizing over integer electric-flux
sectors.  At \(\theta=0\) the leading fractional-probe string tension is
\begin{equation}
  \sigma(Q)
  =
  \kappa M
  \left(1-\cos\frac{2\pi Q}{e}\right)
  +O(M^2).
  \label{eq:massive-schwinger-string-tension}
\end{equation}
It vanishes for \(Q/e\in\mathbb Z\) and is positive, to this order, for
nonintegral \(Q/e\).
\end{proposition}

\begin{proof}
The external worldlines impose Gauss-law discontinuities.  Across a probe of
charge \(Q\), the electric flux changes by \(Q\).  In the action, the theta
term contributes
\[
  \int\dd^2x\,\frac{e\theta}{2\pi}E .
\]
Therefore a region in which the background electric flux is shifted by
\(Q\) is equivalently described, in the bosonized scalar branch, by replacing
\(\theta\) by \(\theta+2\pi Q/e\).  This proves
\eqref{eq:schwinger-theta-shift-probe}.

For \(M=0\), the axial shift symmetry makes the vacuum energy independent of
\(\theta\).  To first order in \(M\), ordinary perturbation theory around the
massless Schwinger vacuum gives the vacuum-energy density
\[
  \varepsilon(\theta)
  =
  -\kappa M\cos\theta+O(M^2),
\]
where the same \(\kappa\) appears in the renormalized mass perturbation
\eqref{eq:massive-schwinger-sine-gordon-perturbation}.  The interval between
widely separated probes has energy density
\(\varepsilon(\theta+2\pi Q/e)\), while the exterior vacuum has
\(\varepsilon(\theta)\).  Their difference is
\eqref{eq:massive-schwinger-string-tension-theta}.  At \(\theta=0\) this is
\eqref{eq:massive-schwinger-string-tension}.  If \(Q/e\) is an integer, the
theta shift is \(2\pi\)-periodic and the leading string tension vanishes; if
\(Q/e\notin\mathbb Z\), the shifted interval lies in a different point of the
theta circle and has positive leading energy density at \(\theta=0\).
\end{proof}

This is confinement of fractional test charges, not confinement of the
dynamical unit-charged fermion.  The latter belongs to the charge lattice
that can screen integral probes.

\section{Mechanisms Isolated by the Exact Solution}

The model is useful because several mechanisms that are intertwined in
higher-dimensional gauge theories become separately computable:
\[
\begin{array}{c|c}
  \hbox{object} & \hbox{exact statement in the Schwinger model} \\ \hline
  \hbox{local anomaly} &
  \partial_\mu J_A^\mu=(e/\pi)E \\
  \hbox{local electric-field spectrum} &
  E\hbox{ creates a scalar of }m_{\rm Sch}^2=e^2/\pi \\
  \hbox{massless probe response} &
  V(R)=Q^2(1-\ee^{-m_{\rm Sch}R})/(2m_{\rm Sch}) \\
  \hbox{mass perturbation} &
  M\bar\psi\psi\leftrightarrow\kappa M\cos(2\sqrt\pi\varphi+\theta) \\
  \hbox{fractional-probe confinement} &
  \sigma(Q)=\kappa M(1-\cos(2\pi Q/e))+O(M^2)
\end{array}
\]
Each line refers to a specified gauge-invariant local observable or to an
explicit external line probe.  This separation is essential: a massive
electric-field two-point function, a screened Wilson-line potential, and a
linear potential for fractional probes in the massive deformation are
different physical statements, even though in this exactly solvable theory
they are controlled by the same current algebra and anomaly.
