\chapter{Wilsonian Effective Actions and Exact Cutoff Flow}
\label{ch:vol2-wilsonian-effective-actions}
\ConventionsEuclidean


\section{Smooth Cutoffs and Effective Actions}

We work in Euclidean signature on \(\R^D\), first for one real scalar field.
Fourier transforms are normalized by
\[
  \phi(x)=\int\frac{\dd^Dk}{(2\pi)^D}\,
  \ee^{\ii k\cdot x}\widetilde\phi(k),
  \qquad
  \widetilde\phi(k)=\int\dd^Dx\,\ee^{-\ii k\cdot x}\phi(x).
\]
For compactness, write
\[
  \int_k F(k):=\int\frac{\dd^Dk}{(2\pi)^D}\,F(k),
  \qquad
  \frac{\delta\widetilde\phi(p)}{\delta\widetilde\phi(k)}
  =(2\pi)^D\delta^{(D)}(p-k).
\]

Gauge theories require extra finite-cutoff data.  A Wilsonian construction for
ordinary scalar or matter variables may be written as a covariance split and a
Gaussian pushforward.  For gauge variables, the corresponding construction is
made inside one of two gauge-compatible frameworks: a finite lattice
formulation with group-valued link variables and gauge-invariant blocking
maps, or a finite-cutoff BV formulation whose action, half-density, bracket,
Laplacian, integration cycle, and coarse-graining map satisfy the quantum
master equation.  The scalar formulas in this chapter supply the ordinary
field model.  Their gauge-theory use is routed through the lattice chapter and
the BV master-formalism chapter, where the gauge quotient, ghosts,
antifields, and Ward identities are part of the finite-cutoff datum.

Let \(\chi:[0,\infty)\to[0,1]\) be a smooth nonincreasing cutoff profile with
\(\chi(0)=1\) and rapid decay as \(t\to\infty\).  The generic
Wilson-Polchinski algebra below only needs the covariance difference
\(C_\Lambda-C_{\Lambda'}\) to be a positive covariance.  Exact preservation of
a nonzero low-source sector by a source-independent Wilsonian action requires
an additional plateau condition, stated later in
\eqref{eq:admissible-plateau-cutoff}; without such a plateau one must keep
source-dependent Wilsonian vertices or prove an explicit leakage estimate.
At cutoff scale \(\Lambda>0\), define
\[
  \chi_\Lambda(k):=\chi(k^2/\Lambda^2),
  \qquad
  C_\Lambda(k):=\frac{\chi_\Lambda(k)}{k^2+m^2}.
\]
Here \(m^2\geq0\) is the mass parameter in the quadratic kernel.  The
Wilsonian action at scale \(\Lambda\) is written
\begin{equation}
  S_\Lambda[\phi]
  =
  S_{\mathrm{kin},\Lambda}[\phi]+L_\Lambda[\phi],
  \label{eq:wilsonian-action-split}
\end{equation}
where
\begin{equation}
  S_{\mathrm{kin},\Lambda}[\phi]
  =
  \frac12\int_k
  \widetilde\phi(-k)\,
  \frac{k^2+m^2}{\chi_\Lambda(k)}\,
  \widetilde\phi(k).
  \label{eq:smooth-cutoff-kinetic-term}
\end{equation}
Perturbation theory around this Gaussian measure uses \(C_\Lambda(k)\) as
propagator.  The cutoff is therefore part of the definition of the covariance,
so every propagator in the Gaussian expansion carries the same regulated
kernel.

Figure~\ref{fig:wilsonian-smooth-cutoff} fixes the cutoff convention used in
this chapter: the smooth profile \(\chi_\Lambda(k)\) is built into the Gaussian
covariance, hence every perturbative propagator carries the same ultraviolet
profile.

\begin{figure}[H]
  \centering
  \resizebox{0.98\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small] at (2.2,2.82) {smooth cutoff profile};
      \draw[-{Latex[length=2mm]}] (-0.15,0)--(4.4,0)
        node[below] {\(t=k^2/\Lambda^2\)};
      \draw[-{Latex[length=2mm]}] (0,-0.15)--(0,2.25)
        node[left] {\(\chi(t)\)};
      \draw[dashed,gray] (1.65,0)--(1.65,1.55);
      \draw[dashed,gray] (0,1.55)--(1.65,1.55);
      \node[below] at (1.65,-0.03) {\(1\)};
      \node[left] at (-0.03,1.55) {\(1\)};
      \draw[blue!65!black,very thick]
        (0,1.55)--(1.28,1.55)
        .. controls (1.55,1.54) and (1.62,1.15) .. (1.72,0.62)
        .. controls (1.84,0.10) and (2.45,0.04) .. (4.0,0.04);
      \node[align=center,blue!65!black] at (1.35,2.02)
        {low momenta\\unchanged};
      \node[align=center,blue!65!black] at (3.2,0.72)
        {rapid\\suppression};
    \end{scope}

    \begin{scope}[shift={(5.35,0)}]
      \node[font=\small] at (2.2,2.82) {lowering the cutoff};
      \draw[-{Latex[length=2mm]}] (-0.15,0)--(4.45,0)
        node[below] {\(|k|\)};
      \draw[-{Latex[length=2mm]}] (0,-0.15)--(0,2.25);
      \fill[orange!35,opacity=0.50]
        (1.95,0.04) rectangle (2.85,1.48);
      \draw[green!50!black,very thick]
        (0,1.55)--(1.65,1.55)
        .. controls (1.88,1.53) and (1.96,1.12) .. (2.06,0.62)
        .. controls (2.18,0.10) and (2.78,0.04) .. (4.05,0.04);
      \draw[orange!85!black,very thick]
        (0,1.55)--(2.55,1.55)
        .. controls (2.78,1.53) and (2.86,1.12) .. (2.96,0.62)
        .. controls (3.08,0.10) and (3.62,0.04) .. (4.05,0.04);
      \draw[dashed] (1.95,0)--(1.95,1.22);
      \draw[dashed] (2.85,0)--(2.85,1.22);
      \node[below] at (1.95,-0.03) {\(\Lambda'\)};
      \node[below] at (2.85,-0.03) {\(\Lambda\)};
      \node[green!50!black] at (1.0,1.25) {\(\chi_{\Lambda'}\)};
      \node[orange!85!black] at (3.45,1.12) {\(\chi_\Lambda\)};
      \node[align=center,fill=white,inner sep=2pt] at (2.45,1.86)
        {shell covariance\\\(C_\Lambda-C_{\Lambda'}\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The plateau cutoff profile drawn here equals one for sufficiently
  small momentum and suppresses modes above \(\Lambda\).  When the cutoff is
  lowered to \(\Lambda'\), the difference of covariances is concentrated near
  the momentum shell between the two scales.  Non-plateau smooth profiles do
  not give an exact low-source identity without adding source-dependent
  Wilsonian vertices.}
  \label{fig:wilsonian-smooth-cutoff}
\end{figure}

The interaction functional \(L_\Lambda\) is an arbitrary translation-invariant
functional compatible with the imposed symmetries.  In a finite spatial
volume with a finite momentum cutoff, \(L_\Lambda\) is an ordinary smooth or
polynomial function of finitely many Fourier modes.  Its vertex expansion is
\begin{align}
  L_\Lambda[\phi]
  &=
  \sum_{n\geq0}\frac1{n!}
  \int\prod_{i=1}^{n-1}\frac{\dd^Dk_i}{(2\pi)^D}\,
  L_\Lambda^{(n)}(k_1,\ldots,k_{n-1})
  \notag\\
  &\hspace{3.35cm}\times
  \widetilde\phi(k_1)\cdots\widetilde\phi(k_{n-1})
  \widetilde\phi(-k_1-\cdots-k_{n-1}).
  \label{eq:general-wilsonian-interaction}
\end{align}
The coefficient functions \(L_\Lambda^{(n)}\) are symmetric in their \(n\)
external momenta after the momentum-conserving argument is restored.  Locality
of the Wilsonian action means that, for momenta small compared with
\(\Lambda\), these coefficient functions admit an expansion in polynomials of
the external momenta, with the nonpolynomial dependence controlled at the
cutoff scale.  The derivative expansion supplies an infinite set of local
coordinates organized by powers of momentum over \(\Lambda\).

The source-dependent generating functional at cutoff \(\Lambda\) is
\begin{equation}
  Z_\Lambda[J]
  =
  \int\dd\mu_{C_\Lambda}(\phi)\,
  \exp\left[
    -L_\Lambda[\phi]
    -\int\dd^Dx\,J(x)\phi(x)
  \right].
  \label{eq:wilsonian-z-lambda}
\end{equation}
The Gaussian measure \(\dd\mu_{C_\Lambda}\) carries
\(S_{\mathrm{kin},\Lambda}\) as in
\eqref{eq:gaussian-reference-measure-from-density}; with the reference density
convention the same expression is
\[
  Z_{0,\Lambda}^{-1}
  \int D_\Lambda^{\rm ref}\phi\,
  \exp\left[
    -S_{\mathrm{kin},\Lambda}[\phi]
    -L_\Lambda[\phi]
    -\int\dd^Dx\,J(x)\phi(x)
  \right].
\]
The normalization \(Z_{0,\Lambda}^{-1}\) is immaterial for connected
correlators with at least one source derivative.
The source-independent Wilsonian condition for lowering \(\Lambda\) to
\(\Lambda'<\Lambda\) is the equality
\begin{equation}
  Z_{\Lambda'}[J]=Z_\Lambda[J]
  \label{eq:low-momentum-generating-function-preserved}
\end{equation}
for sources \(J\) that annihilate the shell covariance,
\(\widehat C_{\Lambda,\Lambda'}J=0\).  For an orthogonal finite-mode split this
means \(J\) belongs to the dual of the retained low-mode space.  For a smooth
profile it is an exact nonzero condition only when the profile has a plateau
and the source support lies strictly inside that plateau.  If the profile is
non-plateau, such as a Gaussian profile, the equality in
\eqref{eq:low-momentum-generating-function-preserved} is not an exact
statement about the source-independent \(L_{\Lambda'}\); one must either add
source-dependent vertices to the pushforward or state and prove a
profile-dependent leakage estimate.  Under the annihilator condition, the
scale label on \(S_\Lambda\) is a label on a description of the same selected
source functional.

\section{Lowering the Cutoff}

The exact Wilsonian step is most transparent with a finite regulator.  Let
\(E_\Lambda\) be the finite-dimensional real vector space of Fourier modes
retained by the ultraviolet regulator, and let \(\dd\mu_{C_\Lambda}\) be the
centered Gaussian measure with covariance \(C_\Lambda\) on this space.  In the
notation of
Definition~\ref{def:regulated-scalar-integration-conventions}, this chapter has
absorbed the quadratic kinetic term into \(\dd\mu_{C_\Lambda}\); equivalently,
\[
  \dd\mu_{C_\Lambda}(\phi)
  =
  Z_{0,\Lambda}^{-1}
  \exp\{-S_{\rm kin,\Lambda}[\phi]\}\,
  D_\Lambda^{\rm ref}\phi .
\]
The interaction \(L_\Lambda\) is integrated against this Gaussian reference
measure.  If
\[
  C_\Lambda=C_{\Lambda'}+\widehat C_{\Lambda,\Lambda'}
\]
with both summands positive covariances on the regulated vector space, then
the Gaussian convolution identity says
\begin{equation}
  \int_{E_\Lambda}F(\phi)\,\dd\mu_{C_\Lambda}(\phi)
  =
  \int_{E_{\Lambda'}}\int_{\widehat E_{\Lambda,\Lambda'}}
  F(\phi'+\phi_{\mathrm{sh}})\,
  \dd\mu_{\widehat C_{\Lambda,\Lambda'}}(\phi_{\mathrm{sh}})\,
  \dd\mu_{C_{\Lambda'}}(\phi')
  \label{eq:finite-regulator-gaussian-convolution}
\end{equation}
for every integrable regulated function \(F\).  Equation
\eqref{eq:finite-regulator-gaussian-convolution} is a theorem about Gaussian
measures on finite-dimensional vector spaces.  Its proof and its continuum
status are logically separate.

\begin{proposition}[Finite-regulator Gaussian pushforward]
\label{prop:finite-regulator-gaussian-pushforward}
Assume \(E_\Lambda=E_{\Lambda'}\oplus \widehat E_{\Lambda,\Lambda'}\) and that
\(C_{\Lambda'}\) and \(\widehat C_{\Lambda,\Lambda'}\) are positive covariance
operators on the two summands.  Let \(C_\Lambda=C_{\Lambda'}\oplus
\widehat C_{\Lambda,\Lambda'}\).  If \(F\) is integrable with respect to the
centered Gaussian measure \(\mu_{C_\Lambda}\), then
\eqref{eq:finite-regulator-gaussian-convolution} holds.
\end{proposition}

\begin{proof}
Let \(X'\) and \(X_{\rm sh}\) be independent centered Gaussian random variables
with values in \(E_{\Lambda'}\) and \(\widehat E_{\Lambda,\Lambda'}\) and
covariances \(C_{\Lambda'}\) and \(\widehat C_{\Lambda,\Lambda'}\).  For every
linear functional \(\ell\in E_\Lambda^*\),
\[
  \mathbb E\,\ee^{\ii\ell(X'+X_{\rm sh})}
  =
  \exp\!\left\{-\frac12\ell(C_{\Lambda'}\ell)\right\}
  \exp\!\left\{-\frac12\ell(\widehat C_{\Lambda,\Lambda'}\ell)\right\}
  =
  \exp\!\left\{-\frac12\ell(C_\Lambda\ell)\right\}.
\]
Thus \(X'+X_{\rm sh}\) has the centered Gaussian law with covariance
\(C_\Lambda\).  Equality of expectations of every integrable \(F\) gives
\eqref{eq:finite-regulator-gaussian-convolution}.
\end{proof}

\begin{hypothesis}[Continuum interpretation of Wilsonian pushforward]
\label{hyp:wilsonian-pushforward-continuum-interpretation}
Choose a sequence of finite regulators \(N\to\infty\), source spaces
\(\mathcal S_{<}\) for momenta below a fixed physical scale, and a topology
\(\mathcal T\) on Schwinger functionals or on local Wilsonian kernels.  A
nonperturbative continuum interpretation of the Wilsonian pushforward requires
the following additional assertions.

First, the interacting finite-regulator source functionals
\[
  Z_{\Lambda,N}[J]
  =
  \int_{E_{\Lambda,N}}
  \exp\{-L_{\Lambda,N}[\phi]-\langle J,\phi\rangle\}\,
  \dd\mu_{C_{\Lambda,N}}(\phi)
\]
must converge in \(\mathcal T\), after the chosen renormalization conditions
are imposed, for \(J\in\mathcal S_<\).  This is an existence statement for the
interacting integral, logically independent of the Gaussian convolution
identity.

Second, for each pair \(\Lambda'<\Lambda\), the finite-regulator pushforwards
defined by \eqref{eq:finite-regulator-gaussian-convolution} must converge in
the same topology to a limiting map that preserves the limiting low-source
functional.  If the infinitesimal Wilson-Polchinski equation is used, the
finite-regulator vector fields and their solutions must also converge in the
chosen function-space topology.
\end{hypothesis}

The finite-regulator proposition is the theorem used in the algebra below.
Hypothesis~\ref{hyp:wilsonian-pushforward-continuum-interpretation} is a
separate regulator-removal problem.  In a model where this hypothesis remains
unproved, the continuum Wilsonian flow is a controlled formal or perturbative
construction, while a nonperturbative use requires the convergence asserted in
Hypothesis~\ref{hyp:wilsonian-pushforward-continuum-interpretation}.

The algebraic step behind Wilsonian RG is the covariance identity
\begin{equation}
  C_\Lambda(k)
  =
  C_{\Lambda'}(k)+\widehat C_{\Lambda,\Lambda'}(k),
  \qquad
  \widehat C_{\Lambda,\Lambda'}(k)
  :=
  \frac{\chi_\Lambda(k)-\chi_{\Lambda'}(k)}{k^2+m^2}.
  \label{eq:wilsonian-covariance-split}
\end{equation}
The first covariance belongs to a low field \(\phi'\).  The second covariance
belongs to an independent shell field \(\phi_{\mathrm{sh}}\), with kinetic term
\begin{equation}
  \widehat S_{\mathrm{kin}}[\phi_{\mathrm{sh}}]
  =
  \frac12\int_k
  \widetilde\phi_{\mathrm{sh}}(-k)\,
  \frac{k^2+m^2}{\chi_\Lambda(k)-\chi_{\Lambda'}(k)}\,
  \widetilde\phi_{\mathrm{sh}}(k).
  \label{eq:shell-kinetic-term}
\end{equation}
The shell field is denoted \(\phi_{\mathrm{sh}}\).  A tilde denotes
Fourier transformation; thus \(\widetilde\phi_{\mathrm{sh}}(k)\) is the shell
field's Fourier mode, while \(\widetilde\phi(k)\) elsewhere is the Fourier mode
of the active field variable \(\phi\).  The hat in
\(\widehat C_{\Lambda,\Lambda'}\) labels the covariance difference; it is not a
field notation, and the shell field will not be denoted by
a hatted \(\phi\).

Figure~\ref{fig:propagator-split-shell-integration} records the exact propagator
split used in the Wilsonian pushforward: active low modes retain
\(C_{\Lambda'}\), and the shell field is integrated with covariance
\(\widehat C_{\Lambda,\Lambda'}\).

\begin{figure}[H]
  \centering
  \resizebox{0.98\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,2.55) {propagator identity};
    \draw[very thick] (-2.4,1.25)--(-1.2,1.25);
    \node at (-1.82,1.55) {\(C_\Lambda\)};
    \node at (-0.9,1.25) {\(=\)};
    \draw[very thick,green!50!black] (-0.55,1.25)--(0.65,1.25);
    \node[green!50!black] at (0.05,1.55) {\(C_{\Lambda'}\)};
    \node at (0.98,1.25) {\(+\)};
    \draw[very thick,orange!85!black,dashed] (1.35,1.25)--(2.55,1.25);
    \node[orange!85!black] at (1.95,1.55) {\(\widehat C_{\Lambda,\Lambda'}\)};
    \node[align=center,green!50!black] at (0.05,0.42) {low field\\\(\phi'\)};
    \node[align=center,orange!85!black] at (1.95,0.42) {shell field\\\(\phi_{\mathrm{sh}}\)};

    \begin{scope}[shift={(5.10,0)}]
      \node[font=\small] at (1.30,2.55) {shell integration};
      \draw[rounded corners=4pt,thick] (0,0.58) rectangle (2.60,1.68);
      \node at (1.30,1.13) {\(L_\Lambda[\phi'+\phi_{\mathrm{sh}}]\)};
      \draw[green!50!black,thick] (0,1.40)--(-0.72,1.82);
      \draw[green!50!black,thick] (0,0.82)--(-0.72,0.42);
      \draw[green!50!black,thick] (2.60,1.40)--(3.32,1.82);
      \draw[green!50!black,thick] (2.60,0.82)--(3.32,0.42);
      \draw[orange!85!black,thick,dashed]
        (2.05,1.60) .. controls (3.05,1.98) and (3.05,0.32) .. (2.05,0.68);
      \node[green!50!black] at (1.30,2.03) {external low modes};
      \node[orange!85!black,align=center] at (3.62,1.34)
        {contract\\shell modes};
      \draw[-{Latex[length=2mm]},thick] (4.12,1.05)--(4.85,1.05);
    \end{scope}

    \begin{scope}[shift={(10.85,0)}]
      \node[font=\small] at (0,2.55) {new interaction};
      \draw[rounded corners=4pt,thick,green!50!black] (-1.30,0.58) rectangle (1.30,1.68);
      \node[green!50!black] at (0,1.13) {\(L_{\Lambda'}[\phi']\)};
      \draw[green!50!black,thick] (-1.30,1.40)--(-2.02,1.82);
      \draw[green!50!black,thick] (-1.30,0.82)--(-2.02,0.42);
      \draw[green!50!black,thick] (1.30,1.40)--(2.02,1.82);
      \draw[green!50!black,thick] (1.30,0.82)--(2.02,0.42);
      \node[align=center,green!50!black] at (0,0.02)
        {all remaining legs\\carry low momenta};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The old Gaussian covariance can be realized as the covariance of
  two independent fields.  Integrating the shell field contracts the dashed
  propagators and produces a new interaction for the low field.}
  \label{fig:propagator-split-shell-integration}
\end{figure}

Before the support of the source is used, the covariance split rewrites the
cutoff-\(\Lambda\) functional as
\begin{align}
  Z_\Lambda[J]
  &=
  \int[D\phi']\,
  \exp\left[
    -S_{\mathrm{kin},\Lambda'}[\phi']
  \right]
  \int[D\phi_{\mathrm{sh}}]\,
  \exp\left[
    -\widehat S_{\mathrm{kin}}[\phi_{\mathrm{sh}}]
    -L_\Lambda[\phi'+\phi_{\mathrm{sh}}]
  \right]
  \notag\\
  &\hspace{2.0cm}\times
  \exp\left[
    -\int\dd^Dx\,J(x)\bigl(\phi'(x)+\phi_{\mathrm{sh}}(x)\bigr)
  \right].
  \label{eq:z-lambda-after-shell-split}
\end{align}
The condition on \(J\) is now used in its precise regulated meaning: the source
belongs to the annihilator of the shell covariance, so that the Gaussian
random variable \(\langle J,\phi_{\mathrm{sh}}\rangle\) has zero variance.
Equivalently, for an orthogonal mode split the source belongs to the dual of
the low subspace.  For a smooth momentum cutoff this exact condition is
available for a nonzero open low-momentum support only when the profile has a
plateau; a non-plateau profile requires the source-dependent construction
introduced in \eqref{eq:source-dependent-wilsonian-action}.  Under the exact
annihilator condition, the source term in
\eqref{eq:z-lambda-after-shell-split} couples only to \(\phi'\).  The shell
field may then be integrated out by the defining identity
\begin{equation}
  \exp\{-L_{\Lambda'}[\phi']\}
  =
  \int[D\phi_{\mathrm{sh}}]\,
  \exp\left[
    -\widehat S_{\mathrm{kin}}[\phi_{\mathrm{sh}}]
    -L_\Lambda[\phi'+\phi_{\mathrm{sh}}]
  \right],
  \label{eq:define-lambda-prime-interaction}
\end{equation}
up to a \(\phi'\)-independent normalization.  Combining
\eqref{eq:shell-kinetic-term} and
\eqref{eq:define-lambda-prime-interaction} with the Gaussian measure for
\(\phi'\) gives \eqref{eq:low-momentum-generating-function-preserved}.  The
definition is exact for each finite regulator by
Proposition~\ref{prop:finite-regulator-gaussian-pushforward}.  Perturbation
theory expands the logarithm of the shell expectation in connected shell
contractions.  The continuum claim is the additional
Hypothesis~\ref{hyp:wilsonian-pushforward-continuum-interpretation}: after the
renormalization conditions are imposed, both the interacting source functional
and the Wilsonian pushforward or its infinitesimal flow must have the stated
regulator-removal limit.

\section{The Wilson-Polchinski Equation}

Let \(\Lambda'\) approach \(\Lambda\).  The shell covariance becomes
infinitesimal, and the definition of \(L_{\Lambda'}\) becomes a differential
equation.  With \(C_\Lambda(k)=\chi_\Lambda(k)/(k^2+m^2)\) and with the
functional-derivative convention already stated, the interaction functional
satisfies
\begin{equation}
  \Lambda\partial_\Lambda L_\Lambda[\phi]
  =
  \frac12\int_k
  \Lambda\partial_\Lambda C_\Lambda(k)
  \left[
    \frac{\delta L_\Lambda}{\delta\widetilde\phi(k)}
    \frac{\delta L_\Lambda}{\delta\widetilde\phi(-k)}
    -
    \frac{\delta^2 L_\Lambda}
      {\delta\widetilde\phi(k)\delta\widetilde\phi(-k)}
  \right].
  \label{eq:wilson-polchinski-flow}
\end{equation}
In \eqref{eq:wilson-polchinski-flow}, \(\widetilde\phi(k)\) denotes the Fourier
mode of the remaining field.  The notation matches the functional derivative
convention; after the infinitesimal shell has been integrated, there is again
only one field argument of \(L_\Lambda\).

The analytic meaning of \eqref{eq:wilson-polchinski-flow} is fixed by the
following status convention.

\begin{definition}[Function spaces for the Wilson-Polchinski equation]
The Wilson-Polchinski equation will be used in three nested settings.

\emph{Finite regulator.}  Put the theory in finite volume and impose a finite
ultraviolet mode cutoff \(\Lambda_{\rm UV}\).  All fields then belong to a
fixed finite-dimensional real vector space \(E_{\Lambda_{\rm UV}}\).  In this
setting \(L_\Lambda\) is a \(C^2\) function on
\(E_{\Lambda_{\rm UV}}\), the functional derivatives in
\eqref{eq:wilson-polchinski-flow} are ordinary derivatives, and
\(\Lambda\partial_\Lambda L_\Lambda\) is the derivative of a \(C^1\) curve in
the topology of uniform convergence on bounded subsets, with two field
derivatives continuous in the same sense.  This is the status of the Gaussian
convolution derivation.

\emph{Formal local expansion.}  Write \(L_\Lambda\) by the vertex expansion
\eqref{eq:general-wilsonian-interaction}.  For each \(n\), each number
\(q\) of momentum derivatives, and each compact set \(K\) in the
momentum-conserving external-momentum domain, define the seminorm
\[
  p_{n,q,K}(L_\Lambda)
  =
  \max_{|\alpha|\leq q}
  \sup_{(k_1,\ldots,k_{n-1})\in K}
  \left|
    \partial^\alpha L_\Lambda^{(n)}(k_1,\ldots,k_{n-1})
  \right|.
\]
The Fr\'echet topology generated by these seminorms is the coefficientwise
smooth topology on vertex kernels.  A formal perturbative solution is a curve
whose vertex kernels satisfy the coefficient equations obtained from
\eqref{eq:wilson-polchinski-flow} order by order in the chosen coupling,
field-number, loop, and derivative expansion.

\emph{Banach RG chart.}  For theorem-level estimates one must choose a norm.
One convenient local model fixes integers \(m\geq0\), a weight \(a>0\), and a
constant \(c>1\), and takes \(\mathcal B_{\Lambda}^{m,a}\) to be the space of
symmetric momentum-conserving kernels for which
\[
  \|L_\Lambda\|_{\Lambda,m,a}
  =
  \sum_{n\geq0} a^n
  \max_{|\alpha|\leq m}
  \sup_{|k_i|\leq c\Lambda}
  \Lambda^{d_n+|\alpha|}
  \left|
    \partial^\alpha L_\Lambda^{(n)}(k_1,\ldots,k_{n-1})
  \right|
\]
is finite.  The constants \(d_n\) encode the chosen engineering dimensions of
the \(n\)-point vertices, and may be changed by a finite reparametrization of
coordinates.  A Banach-space solution on a scale interval \(I\) is a
\(C^1\) curve \(\Lambda\mapsto L_\Lambda\in\mathcal B_{\Lambda}^{m,a}\)
after identifying nearby \(\mathcal B_{\Lambda}^{m,a}\) by dimensionless
momenta \(p_i=k_i/\Lambda\), with
\eqref{eq:wilson-polchinski-flow} holding in the Banach norm.  Existence,
uniqueness, and cutoff-uniform bounds are analytic assertions about this
chosen normed space.
\end{definition}

Since
\[
  \Lambda\partial_\Lambda C_\Lambda(k)
  =
  \frac{\Lambda\partial_\Lambda\chi(k^2/\Lambda^2)}{k^2+m^2},
\]
the integration kernel is supported where the cutoff profile changes, namely
near \(|k|\simeq\Lambda\) for a rapidly decreasing smooth cutoff.  The two
terms in \eqref{eq:wilson-polchinski-flow} have a direct graphical meaning:
the product of first derivatives joins two vertices by one shell propagator,
whereas the second derivative contracts two legs of a single vertex.

Figure~\ref{fig:wilson-polchinski-diagrams} is the graphical form of
\eqref{eq:wilson-polchinski-flow}: one term joins two Wilsonian vertices by a
shell line, and the other closes two legs on one vertex.

\begin{figure}[H]
  \centering
  \resizebox{0.95\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,2.55)
      {\(\Lambda\partial_\Lambda L_\Lambda\)};
    \node at (1.55,1.15) {\(=\)};

    \begin{scope}[shift={(3.25,0)}]
      \node[font=\small] at (1.35,2.55) {join two vertices};
      \draw[rounded corners=3pt,thick] (0.0,0.72) rectangle (0.58,1.28);
      \draw[rounded corners=3pt,thick] (2.10,0.72) rectangle (2.68,1.28);
      \draw[orange!85!black,very thick,dashed] (0.58,1.0)--(2.10,1.0);
      \foreach \a in {130,200,250} {
        \draw[green!50!black,thick] (0.0,1.0)--+(\a:0.78);
      }
      \foreach \a in {-70,20,70} {
        \draw[green!50!black,thick] (2.68,1.0)--+(\a:0.78);
      }
      \node[align=center] at (1.35,0.16)
        {\(\displaystyle
        \frac{\delta L_\Lambda}{\delta\widetilde\phi(k)}
        \frac{\delta L_\Lambda}{\delta\widetilde\phi(-k)}\)};
    \end{scope}

    \node at (7.05,1.15) {\(-\)};

    \begin{scope}[shift={(8.25,0)}]
      \node[font=\small] at (1.25,2.55) {contract one vertex};
      \draw[rounded corners=3pt,thick] (0.75,0.72) rectangle (1.35,1.28);
      \draw[orange!85!black,very thick,dashed]
        (0.75,1.08) .. controls (-0.05,1.85) and (2.10,1.85) .. (1.35,1.08);
      \draw[green!50!black,thick] (0.75,1.18)--(0.18,1.55);
      \draw[green!50!black,thick] (0.75,0.86)--(0.18,0.52);
      \draw[green!50!black,thick] (1.35,1.18)--(1.92,1.55);
      \draw[green!50!black,thick] (1.35,0.86)--(1.92,0.52);
      \node[align=center,fill=white,inner sep=1pt] at (1.25,0.12)
        {\(\displaystyle
        \frac{\delta^2 L_\Lambda}
        {\delta\widetilde\phi(k)\delta\widetilde\phi(-k)}\)};
    \end{scope}

    \node[orange!85!black,align=center,fill=white,inner sep=1pt] at (5.8,2.05)
      {dashed shell line weighted by\\\(\Lambda\partial_\Lambda C_\Lambda(k)\)};
    \node[green!50!black] at (5.8,-0.55) {solid external legs carry momenta below the cutoff};
  \end{tikzpicture}
  }
  \caption{The exact cutoff flow has two local operations.  The first term
  connects two interaction vertices with an infinitesimal shell propagator.
  The second term contracts two legs belonging to the same interaction
  vertex.}
  \label{fig:wilson-polchinski-diagrams}
\end{figure}

In the finite-regulator setting, equation
\eqref{eq:wilson-polchinski-flow} is a closed ordinary differential equation
on \(L_\Lambda\).  In the formal local expansion it becomes an infinite
coupled system of equations for the coefficient kernels and their derivative
expansion.  In a Banach RG chart it becomes a vector-field equation only after
one verifies that the right-hand side maps the chosen open subset of
\(\mathcal B_{\Lambda}^{m,a}\) into the appropriate tangent space and satisfies
the estimates needed for the claimed solution class.  The use of a smooth
cutoff is important: it keeps the infinitesimal shell operation compatible
with a derivative expansion at external momenta small compared with
\(\Lambda\).

The sign in \eqref{eq:wilson-polchinski-flow} is fixed by the direction of the
cutoff derivative.  For \(\Lambda'=\Lambda-\delta\Lambda\), the shell covariance
is
\[
  \delta C(k)
  =
  C_\Lambda(k)-C_{\Lambda'}(k)
  =
  \delta\Lambda\,\partial_\Lambda C_\Lambda(k)+O(\delta\Lambda^2).
\]
The Gaussian convolution formula gives, to first order in \(\delta\Lambda\),
\[
  \ee^{-L_{\Lambda-\delta\Lambda}[\phi]}
  =
  \left[
    1+
    \frac12\int_k
    \delta\Lambda\,\partial_\Lambda C_\Lambda(k)
    \frac{\delta^2}
      {\delta\widetilde\phi(k)\delta\widetilde\phi(-k)}
  \right]
  \ee^{-L_\Lambda[\phi]} .
\]
Since
\[
  \frac{\delta^2\ee^{-L_\Lambda}}
    {\delta\widetilde\phi(k)\delta\widetilde\phi(-k)}
  =
  \left[
    \frac{\delta L_\Lambda}{\delta\widetilde\phi(k)}
    \frac{\delta L_\Lambda}{\delta\widetilde\phi(-k)}
    -
    \frac{\delta^2L_\Lambda}
      {\delta\widetilde\phi(k)\delta\widetilde\phi(-k)}
  \right]\ee^{-L_\Lambda},
\]
and
\[
  L_{\Lambda-\delta\Lambda}
  =
  L_\Lambda-\delta\Lambda\,\partial_\Lambda L_\Lambda+O(\delta\Lambda^2),
\]
the displayed Wilson-Polchinski equation follows.  This derivation is an
identity for the regulated finite-dimensional Gaussian integral; the continuum
notation suppresses the regulator that makes the functional derivatives
ordinary derivatives on a finite vector space.
The \(\phi\)-independent normalization omitted in
\eqref{eq:define-lambda-prime-interaction} is the vacuum-energy coordinate.
If it is retained, it replaces \(L_\Lambda[\phi]\) by
\(L_\Lambda[\phi]+c_\Lambda\) and contributes only
\(\Lambda\partial_\Lambda c_\Lambda\) to the left-hand side of the flow.  The
functional derivatives in the right-hand side annihilate \(c_\Lambda\).  The
displayed equation is therefore the Wilsonian interaction flow modulo this
field-independent normalization, whose value may be fixed separately without
changing any non-vacuum vertex or connected normalized correlator.

\section{Effective Average Action and the Wetterich Equation}
\label{sec:effective-average-action-wetterich}

The Wilsonian interaction \(L_\Lambda\) is obtained by integrating a momentum
shell and leaving an action for the remaining field variable.  There is a
second exact cutoff construction, logically different but derived from the
same regulated integral, in which all modes are integrated at once while an
infrared quadratic term suppresses low-momentum fluctuations.  The resulting
object is a cutoff-dependent 1PI functional.  This section uses the name
\emph{effective average action} for that modified Legendre transform.

Because the earlier source convention in
\eqref{eq:wilsonian-z-lambda} used \(-\langle J,\phi\rangle\), introduce in
this section the source coordinate
\[
  \mathcal J:=-J,
  \qquad
  \langle \mathcal J,\phi\rangle
  :=
  \int\dd^Dx\,\mathcal J(x)\phi(x)
  =
  \int_p \widetilde{\mathcal J}(-p)\widetilde\phi(p),
\]
where \(\int_p\) has the same Fourier normalization as \(\int_k\).  This is a
change of source coordinate only; no physics is attached to the sign.

\begin{definition}[Infrared regulator and effective average action]
\label{def:effective-average-action}
Work first at finite volume and with a finite ultraviolet mode cutoff, so the
real field space is a finite-dimensional vector space \(E_N\).  Let
\(S_N[\phi]\) be the full regulated Euclidean action on \(E_N\), including the
quadratic part and all interaction terms allowed by the chosen regulator, and
let \(D_N\phi\) denote the chosen finite-dimensional reference density.  An
infrared regulator at scale \(\kappa>0\) is a \(C^1\) family of symmetric
nonnegative bilinear forms \(R_{\kappa,N}:E_N\to E_N^*\).  In a
translation-invariant scalar regulator one writes
\begin{equation}
  \Delta S_{\kappa,N}[\phi]
  =
  \frac12\int_p
  \widetilde\phi(-p) R_\kappa(p)\widetilde\phi(p),
  \qquad
  R_\kappa(p)=Z_\kappa\kappa^2 r(p^2/\kappa^2),
  \label{eq:eaa-ir-regulator}
\end{equation}
with \(r(s)>0\) for \(s\) near \(0\) and \(r(s)\to0\) rapidly as
\(s\to\infty\).  The displayed profile is a standard scalar model of the
definition; the finite-regulator derivation below only needs a symmetric
\(C^1\) nonnegative family \(R_{\kappa,N}\).

Define
\begin{align}
  Z_{\kappa,N}[\mathcal J]
  &:=
  \int_{E_N}D_N\phi\,
  \exp\{-S_N[\phi]-\Delta S_{\kappa,N}[\phi]
       +\langle \mathcal J,\phi\rangle\},
  \notag\\
  W_{\kappa,N}[\mathcal J]&:=\log Z_{\kappa,N}[\mathcal J],
  \qquad
  \varphi_{\mathcal J}
  :=
  \frac{\delta W_{\kappa,N}}{\delta\mathcal J}.
  \label{eq:eaa-z-w-average-field}
\end{align}
On a source neighborhood where
\(\mathcal J\mapsto\varphi_{\mathcal J}\) is locally invertible, let
\(\mathcal J_{\kappa,N}[\varphi]\) denote the inverse map.  The effective
average action is the modified Legendre transform
\begin{equation}
  \Gamma_{\kappa,N}[\varphi]
  =
  \langle \mathcal J_{\kappa,N}[\varphi],\varphi\rangle
  -W_{\kappa,N}[\mathcal J_{\kappa,N}[\varphi]]
  -\Delta S_{\kappa,N}[\varphi].
  \label{eq:effective-average-action-definition}
\end{equation}
Equivalently, \(\Gamma_{\kappa,N}+\Delta S_{\kappa,N}\) is the ordinary
Legendre transform of \(W_{\kappa,N}\).  If the integral is defined by a
positive measure, \(W_{\kappa,N}\) is convex in \(\mathcal J\), and
\(\Gamma_{\kappa,N}+\Delta S_{\kappa,N}\) is convex in \(\varphi\).  The
subtracted functional \(\Gamma_{\kappa,N}\) need not itself be convex at
positive \(\kappa\).
\end{definition}

The cutoff \(\kappa\) in this section is an infrared scale.  It should not be
identified with the Wilsonian ultraviolet cutoff \(\Lambda\) unless a
separate matching map has been specified.  At large \(\kappa\), the extra
quadratic term suppresses fluctuations with \(p^2\lesssim \kappa^2\).  At
\(\kappa=0\), if \(R_\kappa\to0\) and the limit exists, the modified Legendre
transform becomes the ordinary 1PI effective action of the regulated theory.

\begin{lemma}[Legendre Hessian identity]
\label{lem:eaa-hessian-identity}
In the finite-regulator setting of
Definition~\ref{def:effective-average-action}, assume that the source-field
map is locally invertible at \(\mathcal J\).  Then
\begin{equation}
  \frac{\delta\Gamma_{\kappa,N}}{\delta\varphi}
  =
  \mathcal J_{\kappa,N}[\varphi]-R_{\kappa,N}\varphi,
  \qquad
  \Gamma_{\kappa,N}^{(2)}[\varphi]+R_{\kappa,N}
  =
  \bigl(W_{\kappa,N}^{(2)}[\mathcal J]\bigr)^{-1},
  \label{eq:eaa-hessian-identity}
\end{equation}
where \(W_{\kappa,N}^{(2)}\) is the connected two-point covariance in the
presence of the source and the inverse is the inverse linear map
\(E_N^*\to E_N\) on the chosen source neighborhood.
\end{lemma}

\begin{proof}
Differentiate
\[
  \Gamma_{\kappa,N}[\varphi]+\Delta S_{\kappa,N}[\varphi]
  =
  \langle \mathcal J,\varphi\rangle-W_{\kappa,N}[\mathcal J],
  \qquad
  \varphi=\frac{\delta W_{\kappa,N}}{\delta\mathcal J}.
\]
The terms containing \(\delta\mathcal J\) cancel because
\(\delta W_{\kappa,N}=\langle\varphi,\delta\mathcal J\rangle\).  Hence
\(\delta(\Gamma_{\kappa,N}+\Delta S_{\kappa,N})/\delta\varphi=\mathcal J\),
which gives the first identity after subtracting
\(\delta\Delta S_{\kappa,N}/\delta\varphi=R_{\kappa,N}\varphi\).  A second
variation gives
\[
  \Gamma_{\kappa,N}^{(2)}+R_{\kappa,N}
  =
  \frac{\delta\mathcal J}{\delta\varphi}.
\]
On the other hand
\(\delta\varphi/\delta\mathcal J=W_{\kappa,N}^{(2)}\).  Local invertibility
of the source-field map says precisely that these two derivatives are inverse
finite-dimensional matrices.
\end{proof}

\begin{theorem}[Finite-regulator Wetterich equation]
\label{thm:finite-regulator-wetterich-equation}
Let \(t=\log\kappa\).  Suppose that \(R_{\kappa,N}\) is \(C^1\) in \(t\), the
regulated integral in \eqref{eq:eaa-z-w-average-field} is differentiable in
\(t\), and the source-field map is locally invertible along the curve under
consideration.  Then, at fixed average field \(\varphi\),
\begin{equation}
  \partial_t\Gamma_{\kappa,N}[\varphi]
  =
  \frac12\,\operatorname{Tr}_{E_N}
  \left[
    \bigl(\Gamma_{\kappa,N}^{(2)}[\varphi]+R_{\kappa,N}\bigr)^{-1}
    \partial_t R_{\kappa,N}
  \right].
  \label{eq:wetterich-equation-finite-regulator}
\end{equation}
\end{theorem}

\begin{proof}
Differentiate the modified Legendre transform
\eqref{eq:effective-average-action-definition} with respect to \(t\) while
holding \(\varphi\) fixed.  As in the preceding lemma, the terms containing
\(\partial_t\mathcal J_{\kappa,N}[\varphi]\) cancel.  Thus
\begin{equation}
  \partial_t\Gamma_{\kappa,N}[\varphi]
  =
  -\partial_t W_{\kappa,N}[\mathcal J]\big|_{\mathcal J}
  -\partial_t\Delta S_{\kappa,N}[\varphi].
  \label{eq:eaa-flow-legendre-cancellation}
\end{equation}
The only explicit \(t\)-dependence in
\(Z_{\kappa,N}[\mathcal J]\) is the regulator, so
\[
  \partial_t W_{\kappa,N}[\mathcal J]\big|_{\mathcal J}
  =
  -\frac12
  \left\langle
    \phi\,(\partial_t R_{\kappa,N})\,\phi
  \right\rangle_{\mathcal J,\kappa}^{\rm full}.
\]
Here the bracket denotes the full source-dependent expectation in the
regulated integral.  Decomposing the full second moment into its connected and
mean parts gives
\[
  \left\langle
    \phi\,(\partial_t R_{\kappa,N})\,\phi
  \right\rangle_{\mathcal J,\kappa}^{\rm full}
  =
  \langle\varphi,(\partial_tR_{\kappa,N})\varphi\rangle
  +
  \operatorname{Tr}_{E_N}\!
  \left[
    W_{\kappa,N}^{(2)}[\mathcal J]\,\partial_tR_{\kappa,N}
  \right].
\]
Since
\[
  \partial_t\Delta S_{\kappa,N}[\varphi]
  =
  \frac12
  \langle\varphi,(\partial_tR_{\kappa,N})\varphi\rangle,
\]
the mean-field terms cancel in
\eqref{eq:eaa-flow-legendre-cancellation}.  The result is
\[
  \partial_t\Gamma_{\kappa,N}[\varphi]
  =
  \frac12\operatorname{Tr}_{E_N}
  \left[
    W_{\kappa,N}^{(2)}[\mathcal J]\,\partial_tR_{\kappa,N}
  \right].
\]
Substituting the Hessian identity
\eqref{eq:eaa-hessian-identity} proves
\eqref{eq:wetterich-equation-finite-regulator}.
\end{proof}

After suppressing the finite regulator from the notation, the formal
continuum version is
\begin{equation}
  \partial_t\Gamma_{\kappa}[\varphi]
  =
  \frac12\,\operatorname{Tr}
  \left[
    \bigl(\Gamma_{\kappa}^{(2)}[\varphi]+R_{\kappa}\bigr)^{-1}
    \partial_t R_{\kappa}
  \right],
  \qquad t=\log\kappa .
  \label{eq:wetterich-equation}
\end{equation}
The trace in \eqref{eq:wetterich-equation} means the trace over spacetime,
internal indices, and statistics, with a minus sign for fermionic variables
when the derivation is repeated in the superalgebraic setting.  For the scalar
translation-invariant regulator \eqref{eq:eaa-ir-regulator}, the factor
\(\partial_t R_\kappa(p)\) is concentrated at momenta \(p^2\simeq\kappa^2\)
for a smooth profile.  This makes the right-hand side formally both infrared
and ultraviolet regulated.  A theorem-level continuum use still requires
more than this formal observation: one must specify the regulator-removal
topology, prove convergence of \(W_{\kappa,N}\) or \(\Gamma_{\kappa,N}\), and
prove that the finite-regulator traces converge uniformly on the field
neighborhood being used.

Figure~\ref{fig:wilsonian-vs-effective-average-action} distinguishes the two RG
objects compared here: the Wilsonian action obtained by integrating out high
modes and the effective average action obtained by a Legendre transform with an
infrared regulator.

\begin{figure}[H]
  \centering
  \resizebox{0.98\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      box/.style={draw,rounded corners=3pt,line width=0.8pt,
        align=center,text width=2.45cm,minimum height=0.78cm},
      widebox/.style={draw,rounded corners=3pt,line width=0.8pt,
        align=center,text width=3.25cm,minimum height=0.88cm},
      arr/.style={-{Latex[length=2mm]},line width=0.85pt}
    }

    \node[font=\small] at (2.8,3.15) {Wilsonian cutoff action};
    \node[box] (suv) at (0,2.25) {\(S_{\Lambda_0}\)};
    \node[widebox] (shells) at (2.95,2.25)
      {integrate shells\\\(\Lambda_0\downarrow\Lambda\)};
    \node[widebox] (lw) at (6.15,2.25)
      {remaining-field action\\\(L_\Lambda[\phi_<]\)};
    \node[widebox] (wlow) at (9.35,2.25)
      {low-source\\connected functional};
    \draw[arr] (suv)--(shells);
    \draw[arr] (shells)--(lw);
    \draw[arr] (lw)--(wlow);

    \node[font=\small] at (2.8,0.80) {effective average action};
    \node[box] (sn) at (0,-0.10) {\(S_N+\Delta S_\kappa\)};
    \node[widebox] (allmodes) at (2.95,-0.10)
      {integrate all modes\\with IR suppression};
    \node[widebox] (wk) at (6.15,-0.10)
      {\(W_{\kappa,N}[\mathcal J]\)};
    \node[widebox] (gk) at (9.35,-0.10)
      {modified Legendre\\\(\Gamma_{\kappa,N}[\varphi]\)};
    \draw[arr] (sn)--(allmodes);
    \draw[arr] (allmodes)--(wk);
    \draw[arr] (wk)--(gk);

    \draw[arr,dashed,blue!70!black] (lw.south)
      .. controls (5.9,1.25) and (6.2,0.95) .. (wk.north);
    \node[blue!70!black,align=center,fill=white,inner sep=2pt] at (6.35,1.08)
      {comparison needs\\compatible cutoffs};
  \end{tikzpicture}
  }
  \caption{The Wilsonian action and the effective average action are distinct
  cutoff-dependent objects.  The Wilsonian construction integrates shells and
  leaves an action for remaining fields.  The effective average action
  integrates all modes with an infrared regulator and then performs a
  modified Legendre transform.  A relation between the two descriptions
  requires a compatible choice of cutoff kernels and a matching map.}
  \label{fig:wilsonian-vs-effective-average-action}
\end{figure}

\begin{example}[Constant-field projection and the local potential flow]
\label{ex:eaa-local-potential-flow}
Assume a single scalar field and project the effective average action to the
derivative expansion
\begin{equation}
  \Gamma_\kappa[\varphi]
  =
  \int\dd^Dx\,
  \left\{
    \frac12 Z_\kappa\,\partial_\mu\varphi\,\partial_\mu\varphi
    +U_\kappa(\varphi)
  \right\}
  +\text{terms with more derivatives or more fields}.
  \label{eq:eaa-derivative-expansion-ansatz}
\end{equation}
For a constant field \(\varphi(x)=v\), the Hessian in momentum space is
\[
  \Gamma_\kappa^{(2)}(p,q;v)
  =
  (2\pi)^D\delta^{(D)}(p+q)
  \bigl(Z_\kappa p^2+U_\kappa''(v)\bigr)
\]
inside this projection.  Dividing \eqref{eq:wetterich-equation} by the
spacetime volume gives
\begin{equation}
  \partial_t U_\kappa(v)
  =
  \frac12\int_p
  \frac{\partial_t R_\kappa(p)}
       {Z_\kappa p^2+R_\kappa(p)+U_\kappa''(v)}
  \label{eq:eaa-local-potential-flow}
\end{equation}
before any expansion in powers of \(U_\kappa''\).  This is an exact identity
only after projection to the ansatz
\eqref{eq:eaa-derivative-expansion-ansatz}; the omitted derivative and field
operators feed back into the exact flow through
\(\Gamma_\kappa^{(2)}[\varphi]\).

For the illustrative limiting profile
\[
  R_\kappa(p)=Z_\kappa(\kappa^2-p^2)\theta(\kappa^2-p^2)
\]
and with \(\eta_\kappa:=-\partial_t\log Z_\kappa\) set to zero, one obtains
\begin{equation}
  \partial_t U_\kappa(v)
  =
  \frac{v_D\,\kappa^D}
       {1+U_\kappa''(v)/(Z_\kappa\kappa^2)},
  \qquad
  v_D:=
  \frac{1}{D\,2^{D-1}\pi^{D/2}\Gamma(D/2)} .
  \label{eq:eaa-litim-local-potential-flow}
\end{equation}
The step-function regulator is not smooth.  In a proof-oriented treatment it
should be replaced by a smooth family approaching it, or else treated in a
function space where the distributional boundary term is controlled.  Formula
\eqref{eq:eaa-litim-local-potential-flow} is therefore best read as an
efficient projection formula, not as a substitute for the smooth-regulator
finite-dimensional derivation.
\end{example}

\begin{remark}[Relation to the Wilson-Polchinski equation]
\label{rem:eaa-polchinski-relation}
Both \eqref{eq:wilson-polchinski-flow} and
\eqref{eq:wetterich-equation} are exact cutoff identities at finite regulator,
but they differentiate different objects.  The Wilson-Polchinski equation
acts on a Wilsonian interaction before the remaining low fields have been
integrated.  The Wetterich equation acts on an infrared-regulated Legendre
transform after the regulated integral has been performed.  Therefore the
vertices \(L_\Lambda^{(n)}\) and \(\Gamma_\kappa^{(n)}\) are not the same
coordinate functions, and equality of beta functions is meaningful only after
a finite matching map has been chosen.

For compatible cutoff kernels, the formal exact-RG literature constructs a
Legendre-type relation between Polchinski and effective-average-action flows;
one convenient account is in Rosten's review, following the earlier work of
Morris and Wetterich.\footnote{C.~Wetterich, \emph{Phys. Lett. B}
\textbf{301} (1993) 90--94; T.~R.~Morris, \emph{Int. J. Mod. Phys. A}
\textbf{9} (1994) 2411--2450; O.~J.~Rosten, arXiv:1003.1366.}  In this
monograph that relation is used only with its hypotheses displayed: a fixed
finite regulator, compatible quadratic kernels, invertible Legendre map, and
specified continuum or perturbative limit.  It is not a license to identify
arbitrary Wilsonian truncations with arbitrary effective-average-action
truncations.
\end{remark}

\begin{remark}[Gauge theories]
The scalar derivation uses an ordinary positive quadratic infrared regulator.
For gauge theories, such a regulator generally breaks the gauge symmetry of
the gauge-fixed functional integral unless it is embedded into a background
field construction or a BV/BRST formulation with the corresponding modified
Ward or Slavnov--Taylor identity.  Thus a gauge-theory use of
\eqref{eq:wetterich-equation} requires extra data: the gauge fixing, ghost and
antifield conventions, the regulator's transformation law, and the identity
that replaces gauge invariance along the flow.  These conditions are part of
the definition of the regulated gauge-theory problem, not optional checks
after a scalar-like flow has been written down.
\end{remark}

\section{Linearized Flow and Scaling Coordinates}

A fixed point of the Wilsonian flow is a scale-invariant interaction
\(L_*\) after all quantities have been written in dimensionless coordinates.
Let \(t=\log\Lambda\).  Near a fixed point write
\[
  L_\Lambda=L_*+\sum_I g_I(\Lambda)\,\mathcal O_I ,
\]
where \(\mathcal O_I\) are local coordinates on theory space.  The linearized
flow has the form
\[
  \frac{\dd g_I}{\dd t}
  =
  \sum_J M_I{}^J g_J+O(g^2).
\]
Choosing eigenvectors of \(M\) gives scaling coordinates \(u_\alpha\):
\begin{equation}
  \frac{\dd u_\alpha}{\dd t}
  =
  -y_\alpha u_\alpha+O(u^2).
  \label{eq:wilsonian-linearized-eigenflow}
\end{equation}
The number \(y_\alpha\) is the relevance exponent.\footnote{This sign
convention matches the 1PI thermal-coordinate convention used in the
Wilson--Fisher fixed-point calculation.  There the dimensionless temperature
coordinate obeys
\(\dd\tau/\dd\log\mu=-y_t\tau+\cdots\), with
\(y_t=D-\Delta_{\phi^2}>0\).  Both displayed differential equations are written
with respect to an increasing scale variable.  The infrared-oriented variables
\(s_{\rm W}=\log(\Lambda_0/\Lambda)\) and
\(s_{\rm 1PI}=\log(\mu_0/\mu)\) reverse the sign and give
\(\dd u_\alpha/\dd s_{\rm W}=y_\alpha u_\alpha+\cdots\) and
\(\dd\tau/\dd s_{\rm 1PI}=y_t\tau+\cdots\).  The Wilsonian cutoff flow and the
1PI subtraction-scale flow are still different vector fields until the finite
matching map between Wilsonian and 1PI coordinates is specified.}
In the convention \(t=\log\Lambda\), decreasing the cutoff moves toward the
infrared, so \eqref{eq:wilsonian-linearized-eigenflow} implies
\[
  u_\alpha(\Lambda)
  =
  u_\alpha(\Lambda_0)
  \left(\frac{\Lambda}{\Lambda_0}\right)^{-y_\alpha}
  +O(u^2).
\]
A coordinate with \(y_\alpha>0\) grows when \(\Lambda\) is lowered.  It is a
relevant coordinate and must be fixed by a renormalization condition, or
tuned to zero when one asks for an infrared trajectory ending at the fixed
point.  A coordinate with \(y_\alpha<0\) is irrelevant: its value at a fixed
physical scale is attracted to a function of the relevant and marginal
coordinates along a renormalized trajectory, provided the nonlinear
cutoff-removal estimates stated below hold.

A theorem-level codimension statement about a critical surface requires a
smooth local model of theory space and a differentiable RG endpoint map.  The
eigenvalue calculation supplies the candidate relevant tuning directions; the
local smoothness and codimension assertion used in this volume is the
following.

\begin{definition}[Banach RG chart]
Fix an integer \(k\geq1\).  A \(C^k\) Banach RG chart near a fixed point is a
triple \((\mathfrak U,\psi,\Phi_s)\) with the following data.  The set
\(\mathfrak U\) is an open neighborhood of the fixed point in a \(C^k\)
Banach manifold \(\mathfrak T\) of regulated Wilsonian actions, modulo the
redundant field redefinitions and symmetry identifications already fixed in
the chart.  The chart map
\[
  \psi:\mathfrak U\longrightarrow X
\]
takes values in a Banach space \(X\), and
\[
  X=X_{\rm rel}\oplus X_{\rm rest}
\]
is a topological direct sum with finite-dimensional relevant block
\(X_{\rm rel}\cong\mathbb R^r\).  The maps \(\Phi_s\), \(s\geq0\), are the
local RG maps from the ultraviolet scale to a lower scale, with
\(s=\log(\Lambda_0/\Lambda)\), and are \(C^k\) wherever the trajectory stays
inside \(\mathfrak U\).  In coordinates write
\[
  \varphi_s=\psi\circ\Phi_s\circ\psi^{-1},
  \qquad
  \pi_{\rm rel}:X_{\rm rel}\oplus X_{\rm rest}\to X_{\rm rel}
\]
for the coordinate RG map and relevant projection.
\end{definition}

\begin{theorem}[Finite-reference critical surface]
\label{thm:finite-reference-critical-surface}
Let \((\mathfrak U,\psi,\Phi_s)\) be a \(C^k\) Banach RG chart and fix a
finite RG time \(s_R=\log(\Lambda_0/\Lambda_R)>0\).  Let
\[
  E_{s_R}(z)=\pi_{\rm rel}\bigl(\varphi_{s_R}(z)\bigr),
  \qquad
  z=(u,w)\in X_{\rm rel}\oplus X_{\rm rest},
\]
be the endpoint map that extracts the relevant coordinates at the reference
scale \(\Lambda_R\).  Suppose \(E_{s_R}\) is defined and \(C^k\) on an open
neighborhood of a point \(z_*\), that \(E_{s_R}(z_*)=0\), and that the
partial derivative
\[
  D_uE_{s_R}(z_*):X_{\rm rel}\longrightarrow X_{\rm rel}
\]
is an isomorphism.  Then, after possibly shrinking the neighborhood of
\(z_*\), the set
\[
  \mathcal C_{s_R}=\{z:E_{s_R}(z)=0\}
\]
is a \(C^k\) embedded Banach submanifold of codimension \(r\).  In the chosen
coordinates it is the graph
\[
  u=G_{s_R}(w)
\]
for a \(C^k\) map \(G_{s_R}\) defined on an open subset of \(X_{\rm rest}\).
\end{theorem}

\begin{proof}
The hypotheses make \(E_{s_R}(u,w)=0\) an equation in the finite-dimensional
relevant block \(X_{\rm rel}\), with \(C^k\) dependence on the remaining
Banach variable \(w\).  Since \(D_uE_{s_R}(z_*)\) is invertible, the
Banach-space implicit function theorem gives a neighborhood of \(z_*\) and a
unique \(C^k\) map \(G_{s_R}\) such that
\[
  E_{s_R}(u,w)=0
  \quad\Longleftrightarrow\quad
  u=G_{s_R}(w)
\]
inside that neighborhood.  A graph over the closed complemented subspace
\(X_{\rm rest}\) is an embedded \(C^k\) Banach submanifold, and its
codimension is \(\dim X_{\rm rel}=r\).
\end{proof}

The theorem is a finite-reference statement.  The infinite-time critical
surface consisting of trajectories that remain in the fixed-point
neighborhood and converge to the fixed point as \(\Lambda\to0\) requires a
stable- or center-stable-manifold theorem for the chosen Banach RG flow, with
the corresponding spectral-gap, smoothness, and existence hypotheses.  The
monograph will not use the phrase ``codimension \(r\) critical surface'' as a
theorem-level assertion unless the local chart, endpoint map, and
invertibility or stable-manifold hypotheses have been stated.

The scaling dimension of the local operator conjugate to \(u_\alpha\) is
\[
  \Delta_\alpha=D-y_\alpha
\]
when the fixed point is conformal and the coupling multiplies
\(\int\dd^Dx\,\mathcal O_\alpha(x)\).  Thus the Wilsonian eigenvalue problem
and the conformal scaling-operator problem are the same local question seen
from two sides: flow in theory space and dilation on local operators.

\section{Redundant Directions and Field Coordinates}

Theory space is coordinatized by actions, but different coordinates can
describe the same correlation functions.  A local change of field variable
\[
  \phi(x)\longmapsto \phi(x)+\epsilon F[\phi](x)
\]
changes the action, to first order in \(\epsilon\), by
\[
  S[\phi]\longmapsto
  S[\phi]
  +
  \epsilon\int\dd^Dx\,
  \frac{\delta S}{\delta\phi(x)}F[\phi](x)
  -
  \epsilon\,\operatorname{Tr}\frac{\delta F}{\delta\phi}
  +O(\epsilon^2).
\]
The trace term is the local Jacobian contribution in a regulated functional
integral.  Operators of the form
\[
  \int\dd^Dx\,
  \frac{\delta S}{\delta\phi(x)}F[\phi](x)
  \quad
  \text{together with the associated Jacobian}
\]
generate redundant directions.  They change the coordinate representative of
the Wilsonian action and the representatives of local composite operators, but
they leave separated-point correlation functions invariant after the
corresponding field redefinition is applied.

The quotient by redundant directions is important in the linearized RG
problem.  An eigenvector of the flow that is tangent to a field redefinition
is a change of local coordinates on theory space.  Relevant, marginal, and
irrelevant directions that classify continuum limits are therefore physical
directions modulo these redundant tangent vectors, with symmetries and
operator normalizations fixed as part of the coordinate choice.

\section{Renormalized Trajectories and Universality}

Let \(u_r\) denote the relevant coordinates and \(u_i\) the irrelevant
coordinates near a fixed point.  In a Banach RG chart where the necessary graph
estimates have been proved, a renormalized trajectory is represented by a
submanifold of theory space on which the irrelevant coordinates are functions
of the relevant ones:
\[
  u_i=F_i(u_r).
\]
The functions \(F_i\) are determined by the requirement that the trajectory can
be extended toward the ultraviolet while remaining in the basin where the
fixed-point expansion is valid.  Perturbatively, they are found by solving the
irrelevant beta-function equations order by order in the relevant and marginal
coordinates.

Universality is the statement that many different microscopic actions flow to
the same renormalized trajectory at a fixed physical scale.  Suppose two bare
actions at cutoff \(\Lambda_0\) have the same relevant coordinates after
tuning, but differ by an irrelevant coordinate \(u_i(\Lambda_0)\).  Linearized
flow gives
\[
  u_i(\Lambda_R)
  =
  \left(\frac{\Lambda_R}{\Lambda_0}\right)^{-y_i}
  u_i(\Lambda_0)
  +\text{nonlinear terms},
  \qquad
  y_i<0,
\]
for a fixed reference scale \(\Lambda_R\ll\Lambda_0\).  Thus the difference in
any low-momentum coordinate or correlator projector whose linearization is
bounded in this chart is
\[
  O\!\left(
    \|u_i(\Lambda_0)\|
    \left(\frac{\Lambda_R}{\Lambda_0}\right)^{-y_i}
  \right),
\]
after the relevant coordinates and field normalizations have been matched.
The suppression exponent is \(-y_i=\Delta_i-D>0\) for the first omitted
irrelevant operator.  Nonlinear versions of this statement require the
Banach-norm estimates used in
Theorem~\ref{thm:finite-order-bphz-wilsonian-matching} and in the
finite-coordinate continuum-limit theorem below.

This gives a precise meaning to effective field theory.  A finite list of
relevant and marginal coordinates must be measured or tuned.  Irrelevant
operators are still present, and their coefficients encode real short-distance
physics, but their effect on low-energy observables is organized by powers of
the ratio of scales and by the symmetries that determine which operators may
appear.

\section{A Quartic-Sextic Toy Flow}

Consider the massless scalar theory in \(D=4\) and keep only two local
coordinates in the interaction,
\begin{equation}
  L_\Lambda[\phi]
  =
  \int\dd^4x\,
  \bigl(g_4(\Lambda)\phi(x)^4+g_6(\Lambda)\phi(x)^6\bigr).
  \label{eq:quartic-sextic-truncation}
\end{equation}
This is a two-dimensional projection of the full Wilsonian flow.  Numerical
normalizations of \(g_4\) and \(g_6\) are included in the constants \(a\) and
\(b\) below.
The corresponding dimensionless coordinates are
\begin{equation}
  \lambda_4(\Lambda):=g_4(\Lambda),
  \qquad
  \lambda_6(\Lambda):=\Lambda^2g_6(\Lambda).
  \label{eq:lambda-four-six-definitions}
\end{equation}
The canonical term in the second equation is \(2\lambda_6\), because
\(\phi^6\) has coefficient of mass dimension \(-2\) in four dimensions.  A
projected Wilsonian beta function is, by definition,
\[
  \beta_A(\lambda):=\Lambda\frac{\dd\lambda_A}{\dd\Lambda}.
\]
The variables \(\lambda_A\) are coordinates on a chosen cutoff scheme, not
observables by themselves.  To isolate this two-coordinate mechanism, choose local
coordinates in which the small-coupling flow begins as
\begin{equation}
  \Lambda\frac{\dd\lambda_4}{\dd\Lambda}
  =
  -a\lambda_6+\cdots,
  \qquad
  \Lambda\frac{\dd\lambda_6}{\dd\Lambda}
  =
  2\lambda_6+b\lambda_4^2+\cdots ,
  \label{eq:quartic-sextic-beta-functions}
\end{equation}
where \(a\) and \(b\) depend on the cutoff profile and on the normalization of
the two operators.  The ellipses denote terms of higher order in the retained
local coordinates and terms involving the omitted coordinates.  The coefficient
\(2\) is the canonical irrelevant eigenvalue in the convention
\(t=\log\Lambda\): at fixed dimensionful \(g_6\), the dimensionless
\(\lambda_6=\Lambda^2g_6\) decreases when the cutoff is lowered.

Assume \(\lambda_6=O(\lambda_4^2)\) along the trajectory and work to the first
nontrivial order.  Let
\[
  h(\Lambda):=\lambda_6(\Lambda)+\frac b2\lambda_4(\Lambda)^2 .
\]
Using \eqref{eq:quartic-sextic-beta-functions},
\begin{equation}
  \Lambda\frac{\dd}{\dd\Lambda}
  \left(\lambda_6+\frac b2\lambda_4^2\right)
  =
  2\left(\lambda_6+\frac b2\lambda_4^2\right)
  +O(\lambda_4h)+O(\lambda_4^3).
  \label{eq:slaving-combination}
\end{equation}
Thus, over a perturbative range in which the variation of \(\lambda_4\) only
affects this equation at higher order,
\[
  h(\Lambda)
  =
  h(\Lambda_*)\left(\frac{\Lambda}{\Lambda_*}\right)^2
  +O(\lambda_4^3).
\]
Along the flow toward the infrared, \(\Lambda\downarrow0\), the transverse
combination \(h\) is suppressed by the canonical irrelevant power.  The
attracting curve is therefore
\begin{equation}
  \lambda_6(\Lambda)
  =
  -\frac b2\lambda_4(\Lambda)^2+O(\lambda_4^3)
  \label{eq:lambda-six-slaved}
\end{equation}
on the attracting curve.  Substitution into the \(\lambda_4\) equation gives
\begin{equation}
  \Lambda\frac{\dd\lambda_4}{\dd\Lambda}
  =
  \frac{ab}{2}\lambda_4^2+O(\lambda_4^3).
  \label{eq:lambda-four-induced-beta}
\end{equation}
This equation is the induced beta function for the retained marginal
coordinate in the reduced coordinate system.  Comparison with a 1PI
renormalization convention requires an explicitly chosen finite
reparametrization between Wilsonian coordinates and the renormalized coupling
used to label Green functions; in the displayed coordinate system the
irrelevant coordinate has been eliminated by solving its own flow equation.

Figure~\ref{fig:quartic-sextic-ir-flow} displays the reduced two-coordinate flow
used in the example: the irrelevant sextic coordinate is slaved to the quartic
coordinate along the renormalized trajectory.

\begin{figure}[H]
  \centering
  \resizebox{0.84\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \draw[-{Latex[length=2mm]},thick] (-0.15,0)--(6.5,0)
      node[below] {\(\lambda_4\)};
    \draw[-{Latex[length=2mm]},thick] (0,-2.15)--(0,2.35)
      node[left] {\(\lambda_6\)};
    \node[below left] at (0,0) {\((0,0)\)};

    \draw[blue!65!black,very thick,domain=0.05:5.8,samples=100]
      plot (\x,{-0.055*\x*\x});

    \foreach \y/\xtip/\offset in {1.85/5.10/0.18,1.25/4.45/0.13,0.70/3.75/0.09,-0.55/2.95/-0.08,-1.35/2.20/-0.15} {
      \draw[green!45!black,thick,dash pattern=on 5pt off 2.5pt,
        -{Latex[length=2mm]}]
        (5.9,\y) .. controls ({0.5*(5.9+\xtip)},{\y-0.75})
        and ({\xtip+0.35},{-0.055*\xtip*\xtip+0.55*\offset})
        .. (\xtip,{-0.055*\xtip*\xtip+\offset});
    }
    \foreach \x in {1.0,1.8,2.6,3.4,4.2,5.0} {
      \draw[blue!65!black,thick,-{Latex[length=1.8mm]}]
        (\x,{-0.055*\x*\x})
        .. controls ({\x-0.30},{-0.055*\x*\x+0.04}) and ({\x-0.60},{-0.055*(\x-0.6)*(\x-0.6)})
        .. ({\x-0.82},{-0.055*(\x-0.82)*(\x-0.82)});
    }
    \node[green!45!black,align=center] at (4.95,1.85)
      {decreasing \(\Lambda\)\\flows toward the curve};
    \node[blue!65!black,align=center,fill=white,inner sep=3pt] at (4.55,-1.82)
      {\(\lambda_6=-\frac b2\lambda_4^2+O(\lambda_4^3)\)\\attracting curve};
    \node[blue!65!black] at (1.0,0.45) {IR};
  \end{tikzpicture}
  }
  \caption{In the two-coordinate projection, the irrelevant coordinate
  \(\lambda_6\) rapidly approaches a curve determined by \(\lambda_4\).  The
  arrows indicate the direction of decreasing cutoff.  Dashed trajectories
  approach distinct neighborhoods of the solid attracting curve; the projection
  is not meant to show finite-time merging of RG characteristics.  The curve is
  drawn in a normalization with \(b>0\); changing the sign
  convention for the \(\phi^6\) coordinate reflects the vertical axis but not
  the attraction statement.}
  \label{fig:quartic-sextic-ir-flow}
\end{figure}

This example is deliberately modest.  It shows one concrete feature of the
full equation: irrelevant local coordinates are generally generated even if
they are absent in the chosen ultraviolet-regulated action, and near a
controlled trajectory they are determined, order by order, by the coordinates
that remain to be fixed at a reference scale.

\section{Renormalizability as a Continuum Limit}

The Wilsonian flow is a downward flow in cutoff.  If an action
\(S_{\Lambda_0}\) is given at a finite ultraviolet cutoff \(\Lambda_0\), then
\eqref{eq:define-lambda-prime-interaction} or
\eqref{eq:wilson-polchinski-flow} constructs \(S_\Lambda\) for
\(\Lambda<\Lambda_0\).  The finite-cutoff construction therefore supplies a
trajectory below \(\Lambda_0\); a description extending to arbitrarily large
\(\Lambda\) is a continuum-limit datum.

Take a regulated path integral at cutoff \(\Lambda_0\), with interaction
\begin{equation}
  L_{\Lambda_0}[\phi]
  =
  \sum_I g_I^{\mathrm{bare}}(\Lambda_0)\,O_I[\phi],
  \label{eq:bare-action-at-lambda-zero}
\end{equation}
where \(O_I\) are local operators compatible with the symmetries.  After
flowing down to a fixed reference scale \(\Lambda_R\), low-momentum Green
functions can be expressed in terms of the Wilsonian coordinates
\(g_I(\Lambda_R)\).  A continuum limit is a limit
\(\Lambda_0\to\infty\) in which the bare coordinates are chosen as functions
of \(\Lambda_0\) so that selected physical coordinates at \(\Lambda_R\) are
held fixed and the remaining Wilsonian coordinates at \(\Lambda_R\) approach
limits.
The theorem-level statement of this assertion is
Theorem~\ref{thm:conditional-finite-coordinate-wilsonian-continuum-limit}
below.  It is conditional: it assumes the tuning problem, the irrelevant
semigroup estimate, and the convergence of the generated integral in a
specified finite coordinate projection.  The Wilsonian mechanism is the proof
that these hypotheses imply a cutoff-independent reference-scale point, not a
proof that the hypotheses hold in every regulated model.

In the quartic-sextic projection, impose the ultraviolet boundary condition
\[
  \lambda_4(\Lambda_0)=\lambda_4^{\mathrm{bare}}(\Lambda_0),
  \qquad
  \lambda_6(\Lambda_0)=0,
\]
and choose \(\lambda_4^{\mathrm{bare}}(\Lambda_0)\) so that
\[
  \lambda_4(\Lambda_R)=\lambda_4^R
\]
is fixed at the reference scale.  The Wilsonian renormalizability question in
this projection is whether
\begin{equation}
  \lim_{\Lambda_0\to\infty}
  \lambda_6(\Lambda_R)
  \bigm|_{\lambda_4(\Lambda_R)=\lambda_4^R}
  \label{eq:lambda-six-continuum-limit-question}
\end{equation}
exists.  Under perturbative control, with small couplings along the trajectory
and sufficiently mild dependence of the beta functions on the irrelevant
coordinate, the limit can be seen directly from the transverse coordinate
\(h\).  The boundary condition \(\lambda_6(\Lambda_0)=0\) gives
\[
  h(\Lambda_0)
  =
  \frac b2\lambda_4(\Lambda_0)^2 .
\]
After tuning \(\lambda_4(\Lambda_R)=\lambda_4^R\), the homogeneous memory of
this ultraviolet boundary condition is
\[
  h(\Lambda_R)
  =
  h(\Lambda_0)
  \left(\frac{\Lambda_R}{\Lambda_0}\right)^2
  +O\bigl((\lambda_4^R)^3\bigr)
\]
within the perturbative domain.  Therefore
\[
  \lambda_6(\Lambda_R)
  =
  -\frac b2(\lambda_4^R)^2
  +O\bigl((\lambda_4^R)^3\bigr)
  +O\left(\left(\frac{\Lambda_R}{\Lambda_0}\right)^2\right),
\]
where the coefficient of the last term is bounded by the assumed control of
the bare trajectory.  This displays the continuum-limit mechanism: the
irrelevant coordinate at the fixed physical scale is selected by the
renormalized trajectory, while the memory of the microscopic boundary
condition is power-suppressed.

Figure~\ref{fig:continuum-limit-wilsonian} shows the continuum-limit mechanism
in the same two-coordinate model: different microscopic boundary conditions
approach the same finite-scale renormalized trajectory up to
\((\Lambda_R/\Lambda_0)^2\) corrections.

\begin{figure}[H]
  \centering
  \resizebox{0.88\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      midflow/.style={
        postaction={
          decorate,
          decoration={
            markings,
            mark=at position 0.60 with {\arrow{Latex[length=2mm]}}
          }
        }
      }
    }
    \draw[-{Latex[length=2mm]},thick] (-0.2,0)--(8.0,0)
      node[below] {\(\lambda_4\)};
    \draw[-{Latex[length=2mm]},thick] (0,-1.65)--(0,3.05)
      node[left] {\(\lambda_6\)};

    \draw[dashed,green!45!black] (1.35,-1.45)--(1.35,2.65);
    \node[green!45!black,below] at (1.35,-1.45) {\(\lambda_4^R\)};

    \draw[blue!65!black,very thick,domain=0.35:7.35,samples=120]
      plot (\x,{0.32+0.07*(\x-1.35)+0.045*(\x-1.35)*(\x-1.35)});
    \node[blue!65!black,align=center,fill=white,inner sep=2pt] at (5.6,2.38)
      {renormalized\\trajectory curve};

    \foreach \x/\y in {7.2/0,6.4/0,5.6/0,4.8/0,4.0/0} {
      \fill[purple!80!black] (\x,\y) circle (1.8pt);
      \draw[purple!80!black,thick,midflow]
        (\x,\y)
        .. controls ({0.72*\x+0.62},{1.40}) and (2.45,0.55)
        .. (1.35,0.32);
    }
    \node[purple!80!black,align=center,fill=white,inner sep=1pt] at (4.25,1.14)
      {RG flow\\toward \(\Lambda_R\)};
    \node[purple!80!black,align=center] at (5.85,-0.62)
      {UV-regulated actions\\\(\lambda_6(\Lambda_0)=0\)};

    \fill[green!45!black] (1.35,0.32) circle (2.5pt);
    \node[green!45!black,align=left,fill=white,inner sep=1pt] at (2.05,-0.32)
      {\((\lambda_4^R,\lambda_6(\Lambda_R))\)};

    \draw[-{Latex[length=2mm]},red!70!black,thick,dashed]
      (0.22,0.32)--(1.25,0.32);
    \node[red!70!black,align=center,fill=white,inner sep=2pt] at (0.72,0.85)
      {limiting value\\as \(\Lambda_0\to\infty\)};

    \node[gray!70!black] at (6.95,0.38) {\(\Lambda_0\)};
    \node[gray!70!black] at (7.35,0.82) {\(\Lambda_0'\)};
    \draw[-{Latex[length=1.7mm]},gray!70!black] (6.85,0.45)--(7.25,0.72);
    \node[gray!70!black] at (7.55,1.08) {increasing cutoff};
  \end{tikzpicture}
  }
  \caption{For each ultraviolet cutoff \(\Lambda_0\), the bare quartic
  coupling is tuned so that \(\lambda_4(\Lambda_R)=\lambda_4^R\).  The
  continuum-limit statement asks whether the induced value of
  \(\lambda_6(\Lambda_R)\) approaches a limit as \(\Lambda_0\) is removed.}
  \label{fig:continuum-limit-wilsonian}
\end{figure}

The same analysis allows a finite nonzero value of the dimensionless
irrelevant coordinate at the ultraviolet cutoff.  In the example,
\[
  g_6^{\mathrm{bare}}(\Lambda_0)
  =
  \Lambda_0^{-2}\lambda_6(\Lambda_0).
\]
Thus a finite \(\lambda_6(\Lambda_0)\) corresponds to a dimensionful bare
coefficient suppressed by the cutoff.  This is the form produced by
lattice regularizations with lattice spacing \(a=\Lambda_0^{-1}\): many local
operators are present at the cutoff scale, but after tuning the relevant and
marginal coordinates, irrelevant coordinates approach their continuum
trajectory values at fixed physical scale.
For a hypercubic lattice, the edge of the Brillouin zone is
\(\Lambda_{\rm BZ}=\pi/a=\pi\Lambda_0\).  The estimates in this chapter use
\(\Lambda_0=a^{-1}\) as the inverse-spacing cutoff convention; using
\(\Lambda_{\rm BZ}\) instead only rescales the constants multiplying
cutoff-suppressed powers.

The perturbative Wilsonian statement has explicit hypotheses.  Along the
trajectory under consideration, the couplings must remain small enough for the
local-coordinate expansion and the omitted-operator estimates to be valid; the
derivatives such as \(\partial\beta_6/\partial\lambda_6\) and
\(\partial\beta_4/\partial\lambda_4\) must remain controlled; and the beta
function for the retained marginal coordinate must vary slowly enough along
the trajectory that the irrelevant directions are genuinely suppressed.  In
this regime, the linearized irrelevant eigenvalues and the generated-integral
estimate give the power-counting organization of local operators in
perturbative renormalization.

This finite-coordinate estimate should not be mistaken for a constructive
existence theorem for a non-Gaussian four-dimensional scalar continuum limit.
Theorem~\ref{thm:standard-four-dimensional-scalar-triviality} states that,
for the positive lattice \(\lambda\phi^4_4\) and nearest-neighbor
ferromagnetic Ising-type scaling-limit problems covered by the rigorous
triviality results cited there, the critical or near-critical continuum limits
are Gaussian.  The Wilsonian estimate above is a local perturbative
mechanism: it explains how irrelevant coordinates lose memory of their
finite-cutoff boundary values when a controlled trajectory with fixed
renormalized coordinates is assumed.  It does not prove that such a controlled
trajectory exists to arbitrarily high cutoff at nonzero renormalized quartic
coupling.  For the specified standard families and covered critical regimes,
the triviality result says that the limiting connected correlations beyond the
two-point function vanish.

\section{A Local Estimate for Cutoff Removal}

The quartic-sextic example is the one-dimensional irrelevant-coordinate
version of a general local estimate.  This estimate is a finite-coordinate
statement: choose a projection of the Wilsonian action onto local coordinates
\[
  z=(u,v)\in U\times V,
\]
where \(u=(u^1,\ldots,u^r)\) are the coordinates fixed by renormalization
conditions at the reference scale \(\Lambda_R\), and
\(v=(v^1,\ldots,v^s)\) are irrelevant coordinates retained in the projection.
Let
\[
  t=\log\Lambda,\qquad
  t_R=\log\Lambda_R,\qquad
  t_0=\log\Lambda_0 .
\]
In the coordinate neighborhood under consideration, write the projected flow
as
\begin{equation}
  \frac{\dd u^a}{\dd t}=B^a(u,v),
  \qquad
  \frac{\dd v^\rho}{\dd t}
  =
  A^\rho{}_\sigma v^\sigma+F^\rho(u,v).
  \label{eq:finite-coordinate-wilsonian-flow-split}
\end{equation}
The matrix \(A\) is the linear irrelevant block.  In the convention
\(t=\log\Lambda\), an irrelevant coordinate has a positive eigenvalue in
\(A\); the flow is therefore contracting in that coordinate when the cutoff
is lowered from \(t_0\) to \(t_R<t_0\).  We assume a norm on \(V\) and
constants \(K_A,\omega>0\) such that
\begin{equation}
  \|e^{A\tau}\|
  \leq
  K_A e^{\omega\tau},
  \qquad
  \tau\leq0 .
  \label{eq:irrelevant-backward-semigroup-bound}
\end{equation}
Thus \(e^{A(t_R-t_0)}\) is bounded by
\[
  K_A\left(\frac{\Lambda_R}{\Lambda_0}\right)^\omega .
\]

For a solution of \eqref{eq:finite-coordinate-wilsonian-flow-split} on
\([t_R,t_0]\), variation of constants gives
\begin{equation}
  v(t_R)
  =
  e^{A(t_R-t_0)}v(t_0)
  -
  \int_{t_R}^{t_0}
  e^{A(t_R-s)}F(u(s),v(s))\,\dd s .
  \label{eq:irrelevant-coordinate-variation-of-constants}
\end{equation}
The first term is the direct memory of the irrelevant bare coordinate.  The
second term is generated by the flow along the trajectory.  These two terms
have different logical roles: suppression of the first term follows from the
irrelevant eigenvalues, while existence of the cutoff-removal limit requires
control of the generated integral and of the tuning used to hold \(u(t_R)\)
fixed.

\begin{proposition}[Suppression of irrelevant boundary memory]
Assume \eqref{eq:irrelevant-backward-semigroup-bound}.  If the irrelevant bare
coordinate satisfies \(\|v(t_0)\|\leq M\) uniformly in \(t_0\), then the direct
boundary contribution to \(v(t_R)\) obeys
\[
  \|e^{A(t_R-t_0)}v(t_0)\|
  \leq
  K_A M
  \left(\frac{\Lambda_R}{\Lambda_0}\right)^\omega .
\]
If the first omitted irrelevant coordinate has exponent \(\omega_{\rm min}\),
then the same estimate begins with the power
\((\Lambda_R/\Lambda_0)^{\omega_{\rm min}}\).
\end{proposition}

\begin{proof}
Set \(\tau=t_R-t_0\leq0\) in
\eqref{eq:irrelevant-backward-semigroup-bound}.  Since
\(\exp[\omega(t_R-t_0)]=(\Lambda_R/\Lambda_0)^\omega\), the displayed
estimate follows immediately.
\end{proof}

The generated term in
\eqref{eq:irrelevant-coordinate-variation-of-constants} is the part that
selects the continuum graph.  The precise finite-coordinate hypothesis is the
existence, after tuning the retained bare coordinates, of a limit
\begin{equation}
  V^\rho_R(u_R)
  =
  \lim_{\Lambda_0\to\infty}
  \left[
    -
    \int_{t_R}^{t_0}
    \left(e^{A(t_R-s)}F(u(s),v(s))\right)^\rho
    \dd s
  \right],
  \label{eq:continuum-graph-generated-irrelevant-coordinate}
\end{equation}
where the trajectory is constrained by
\[
  u(t_R)=u_R .
\]
When this limit exists uniformly for \(u_R\) in a small coordinate
neighborhood, the finite projection of the renormalized trajectory at
\(\Lambda_R\) is the graph
\[
  v^\rho=V^\rho_R(u_R).
\]
For finite \(\Lambda_0\), define the generated-integral remainder
\[
  R_F(t_0;u_R)
  =
  -
  \int_{t_R}^{t_0}
  e^{A(t_R-s)}F(u(s),v(s))\,\dd s
  -
  V_R(u_R).
\]

\begin{theorem}[Conditional finite-coordinate Wilsonian continuum limit]
\label{thm:conditional-finite-coordinate-wilsonian-continuum-limit}
Fix \(u_R\) and suppose that, for every sufficiently large \(t_0\), the
retained bare coordinates can be tuned so that \(u(t_R)=u_R\).  Assume:
\begin{enumerate}[label=(\roman*)]
  \item the solution remains in a coordinate neighborhood where
  \eqref{eq:finite-coordinate-wilsonian-flow-split} is valid;
  \item the irrelevant semigroup satisfies
  \eqref{eq:irrelevant-backward-semigroup-bound};
  \item the irrelevant bare coordinates are uniformly bounded;
  \item the generated integral in
  \eqref{eq:continuum-graph-generated-irrelevant-coordinate} converges
  uniformly with
  \[
    \|R_F(t_0;u_R)\|
    \leq
    K_F\left(\frac{\Lambda_R}{\Lambda_0}\right)^{\omega_F}
  \]
  for some \(K_F,\omega_F>0\).
\end{enumerate}
For the tuned finite-cutoff trajectory, write
\[
  z_R(t_0;u_R)
  :=
  \bigl(u(t_R;t_0,u_R),\,v(t_R;t_0,u_R)\bigr).
\]
Then the reference-scale Wilsonian coordinates have the limit
\begin{equation}
  \lim_{\Lambda_0\to\infty}z_R(t_0;u_R)
  =
  \bigl(u_R,V_R(u_R)\bigr)
  \label{eq:finite-coordinate-wilsonian-continuum-limit}
\end{equation}
in the retained coordinate space \(U\times V\), and the irrelevant
coordinates at the reference scale obey
\[
  v(t_R)
  =
  V_R(u_R)
  +
  O\!\left[
    \left(\frac{\Lambda_R}{\Lambda_0}\right)^{\min(\omega,\omega_F)}
  \right],
  \qquad
  \Lambda_0\to\infty ,
\]
with the exponent replaced by the smallest retained irrelevant exponent when
several irrelevant coordinates are present.
Consequently, if \(Q\) is any continuous finite-coordinate observable,
projected vertex, or low-momentum kernel functional defined in a neighborhood
of \((u_R,V_R(u_R))\), then
\[
  \lim_{\Lambda_0\to\infty}Q\bigl(z_R(t_0;u_R)\bigr)
  =
  Q\bigl(u_R,V_R(u_R)\bigr).
\]
If \(Q\) is Lipschitz in the chosen coordinate norm, the same power gives an
error bound for this convergence, with a constant multiplied by the Lipschitz
constant of \(Q\).
\end{theorem}

\begin{proof}
Equation \eqref{eq:irrelevant-coordinate-variation-of-constants} writes
\(v(t_R)\) as the sum of the boundary contribution and the generated
contribution.  The preceding proposition bounds the boundary contribution by
\((\Lambda_R/\Lambda_0)^\omega\).  The generated contribution is
\(V_R(u_R)+R_F(t_0;u_R)\), with \(R_F\) bounded by the fourth hypothesis.
Combining the two estimates gives the stated asymptotic form.
The relevant coordinate is fixed by the tuning condition
\(u(t_R;t_0,u_R)=u_R\), so the full retained coordinate vector converges to
\((u_R,V_R(u_R))\).  The final assertion follows by continuity of \(Q\), and
the quantitative estimate follows from the Lipschitz bound applied to the
coordinate difference.
\end{proof}

The hypotheses of
Theorem~\ref{thm:conditional-finite-coordinate-wilsonian-continuum-limit} are
the mathematical content of a
Wilsonian continuum-limit argument in a chosen local projection.  The tuning
condition \(u(t_R)=u_R\) is a statement about the endpoint map from bare
coordinates at \(\Lambda_0\) to renormalized coordinates at \(\Lambda_R\).
In perturbation theory this endpoint map is usually solved order by order.
When the linearized endpoint map in the retained coordinates is invertible,
the implicit function theorem gives the local tuning functions at finite
\(\Lambda_0\).  The remaining task is to prove estimates, uniform in
\(\Lambda_0\), that keep the trajectory in the assumed coordinate
neighborhood and make the generated integral converge.

\begin{remark}[Status of the cutoff-removal hypotheses]
The finite-coordinate continuum-limit theorem is conditional.  It proves
the displayed asymptotic estimate from the four listed hypotheses; it does not
prove those hypotheses for an arbitrary quantum field theory, nor even for an
arbitrary finite projection of a regulated action.  Hypothesis (ii) is a
backward semigroup estimate for the irrelevant part of the linearized
projected RG equation in a chosen norm.  It is immediate in a diagonal
finite-dimensional toy model with strictly negative irrelevant exponents, but
in an actual Wilsonian space it is an analytic estimate about the chosen
Banach or Fr\'echet space of functionals and the linearized cutoff-flow
operator.  Hypothesis (iv) is still stronger: it is the uniform convergence,
after the retained coordinates are tuned, of the generated irrelevant
coordinate.  This is one form of the cutoff-removal problem itself.

Several rigorous results prove analogues of these estimates only after adding
substantial structure.  Polchinski's smooth-cutoff argument proves
perturbative renormalizability order by order for renormalizable Euclidean
Lagrangians by deriving inductive bounds on cutoff-dependent vertices after
the relevant parts are fixed; it is not a nonperturbative construction of an
interacting continuum measure.\footnote{J.~Polchinski, \emph{Nucl. Phys. B}
\textbf{231} (1984) 269--295.}  Salmhofer's rigorous RG framework proves
perturbative flow estimates, especially for fermionic systems, in specified
norms and small-coupling regimes.\footnote{M.~Salmhofer,
\emph{Renormalization: An Introduction}, Springer, 1999.}  The
Brydges--Kennedy tree and Mayer expansion technology supplies iterative bounds
used in constructive RG, but every application still verifies its own
covariance estimates, norms, tuning equations, and convergence
criteria.\footnote{D.~C.~Brydges and T.~Kennedy, \emph{J. Stat. Phys.}
\textbf{48} (1987) 19--49.}

The nonperturbative status is model-dependent.  Constructive scalar field
theory gives non-Gaussian \(\phi^4_2\) and \(\phi^4_3\) continuum theories
under its own ultraviolet, infrared, and stability hypotheses.\footnote{See
J.~Glimm and A.~Jaffe, \emph{Quantum Physics: A Functional Integral Point of
View}, second edition, Springer, 1987, for the constructive scalar-field
framework and references to the original constructions.}  By contrast, the
standard positive four-dimensional scalar and nearest-neighbor Ising-type
families covered by
Theorem~\ref{thm:standard-four-dimensional-scalar-triviality} have Gaussian
scaling limits.  Therefore later use of a Wilsonian continuum-limit estimate
must say which semigroup, tuning, and convergence estimates have been proved
in the chosen framework and which are being imposed as assumptions.
\end{remark}

\begin{remark}[Renormalons and Wilsonian data]
Assumption (iv) is a continuum-limit hypothesis on the Wilsonian generated
integral.  It is stronger than the existence of a formal perturbative expansion
for the projected coordinates.  A renormalon analysis is one way in which
perturbation theory can diagnose the failure of a naive coefficientwise
description to determine the full generated integral.  In a running-coupling
expansion, integration over very small or very large intermediate momenta can
produce factorially growing coefficients and Borel-plane singularities.  When
the corresponding ambiguity has the size of a power correction, the Wilsonian
description must supply the missing datum through retained nonperturbative
coordinates, through an operator-product-expansion matrix element, or through
another matched effective description.  Thus renormalons do not replace the
cutoff-removal hypothesis; they identify particular ways in which that
hypothesis contains information beyond formal perturbation theory.
\end{remark}

Figure~\ref{fig:finite-coordinate-cutoff-removal-estimate} packages the
finite-coordinate estimate used in this discussion: regulator-dependent
coordinates are compared at a fixed physical scale before the ultraviolet
cutoff is removed.

\begin{figure}[H]
  \centering
  \resizebox{0.92\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      box/.style={draw,rounded corners=3pt,line width=0.8pt,
        align=center,text width=2.35cm,minimum height=0.78cm},
      arr/.style={-{Latex[length=2mm]},line width=0.85pt}
    }
    \draw[-{Latex[length=2mm]},thick] (-0.25,0)--(10.2,0)
      node[below] {\(t=\log\Lambda\)};
    \draw[thick] (1.2,-0.12)--(1.2,0.12);
    \draw[thick] (8.8,-0.12)--(8.8,0.12);
    \node[below] at (1.2,-0.12) {\(t_R\)};
    \node[below] at (8.8,-0.12) {\(t_0\)};

    \node[box] (ren) at (1.25,1.05)
      {renormalization\\condition\\\(u(t_R)=u_R\)};
    \node[box] (bare) at (8.75,1.05)
      {bounded\\irrelevant\\bare coordinate\\\(v(t_0)\)};
    \node[box] (gen) at (5.10,2.28)
      {generated integral\\along the tuned\\trajectory};
    \node[box] (graph) at (1.25,3.18)
      {continuum graph\\\(v=V_R(u_R)\)};

    \draw[arr,blue!70!black] (bare.south west)
      .. controls (6.25,0.25) and (3.65,0.25) .. (ren.south east);

    \draw[arr,orange!85!black,dash pattern=on 5pt off 2.5pt] (gen)--(graph);
    \draw[arr,green!45!black,densely dotted] (ren)--(graph);
    \draw[arr,orange!85!black,dash pattern=on 5pt off 2.5pt] (bare.north west)
      .. controls (7.30,2.10) and (6.35,2.28) .. (gen.east);
    \draw[arr,orange!85!black,dash pattern=on 5pt off 2.5pt] (ren.north east)
      .. controls (2.55,1.95) and (3.45,2.20) .. (gen.west);
  \end{tikzpicture}
  }
  \caption{A finite-coordinate Wilsonian continuum-limit estimate separates
  direct irrelevant boundary memory from the generated integral along the
  tuned trajectory.  Solid, dashed, and dotted arrows denote boundary memory,
  generated integral terms, and the renormalization-condition map respectively.
  Irrelevant boundary memory is power-suppressed by the cutoff ratio.  The
  continuum graph at \(\Lambda_R\) is the limiting value of the generated
  contribution, provided the stated uniform estimates hold.}
  \label{fig:finite-coordinate-cutoff-removal-estimate}
\end{figure}

For a lattice regulator with lattice spacing \(a=\Lambda_0^{-1}\), a bounded
dimensionless irrelevant bare coordinate \(v^\rho(t_0)\) corresponds to a
dimensionful coefficient multiplied by the appropriate positive power of
\(a\).  Theorem~\ref{thm:conditional-finite-coordinate-wilsonian-continuum-limit}
explains the resulting universality statement in this local projection: after
the retained coordinates are tuned at
\(\Lambda_R\), bounded changes of microscopic irrelevant lattice couplings
alter \(v(t_R)\) only by powers of \(\Lambda_R/\Lambda_0\), while the limiting
value belongs to the continuum graph selected by the flow.

\section{The Bridge to BPHZ and 1PI Coordinates}

The preceding sections define the Wilsonian action \(S_\Lambda\) by Gaussian
pushforward.  Exact preservation of a source-independent low-source functional
requires the annihilator condition explained above, and for the smooth
source-independent theorem below this will be implemented by an admissible
plateau cutoff.  Earlier chapters defined BPHZ subtractions and 1PI
renormalized coordinates.  These are different constructions, so their
relation must be stated with the maps that connect them.

Fix a finite ultraviolet regulator \(\Lambda_0\), a smooth intermediate cutoff
\(\Lambda\), and a Euclidean subtraction scale \(\mu\) satisfying
\[
  \mu\ll\Lambda\ll\Lambda_0 .
\]
Keep distinct names for the two RG times:
\[
  t_{\rm W}:=\log\Lambda,
  \qquad
  t_{\rm 1PI}:=\log\mu .
\]
They parameterize different vector fields.  The Wilsonian time \(t_{\rm W}\)
acts on the cutoff action before the remaining low field is integrated, while
\(t_{\rm 1PI}\) acts on projected 1PI vertices after the connected functional
has been Legendre transformed.  To compare the two flows, choose a fixed
matching ratio
\[
  \zeta:=\frac{\Lambda}{\mu}\gg 1 .
\]
Along such a matched comparison,
\[
  t_{\rm W}=t_{\rm 1PI}+\log\zeta,
  \qquad
  \left.\frac{\dd}{\dd t_{\rm 1PI}}\right|_{\zeta}
  =
  \left.\frac{\dd}{\dd t_{\rm W}}\right|_{\zeta}.
\]
If \(u_A(\Lambda)\) are Wilsonian coordinates and \(g_I(\mu)\) are 1PI
coordinates, the bridge is a finite matching map at fixed \(\zeta\),
\[
  g_I(\mu)
  =
  M_I(u(\Lambda),\zeta)
  +
  O(\zeta^{-\Delta_{\rm irr}}),
\]
where \(\Delta_{\rm irr}>0\) is the smallest suppression exponent among the
omitted irrelevant coordinates in the chosen local expansion.  Differentiating
at fixed \(\zeta\) gives the corresponding relation between beta functions,
\[
  \beta_I^{\rm 1PI}
  =
  \frac{\partial M_I}{\partial u_A}\,
  \beta_A^{\rm W}
  +
  O(\zeta^{-\Delta_{\rm irr}}),
\]
with summation over \(A\).  Thus the two uses of \(t\) in the preceding
chapters have the same orientation only after this matching convention is
chosen; without the matching map they refer to different mathematical
constructions.

The external momenta used in the 1PI projectors are assumed to have size
\(O(\mu)\) and to be nonexceptional.  Let
\[
  W_{\Lambda_0}[J]=\log Z_{\Lambda_0}[J]
\]
be the connected generating functional of the theory regulated at ultraviolet
scale \(\Lambda_0\), with the cutoff-dependent local terms in the
path-integral Lagrangian chosen so that the selected 1PI coordinates have
finite limits.
Let \(L_\Lambda=L_{\Lambda,0}^{\rm src}\) be the source-independent Wilsonian
interaction obtained from \eqref{eq:define-lambda-prime-interaction} by
integrating modes between \(\Lambda_0\) and \(\Lambda\).  For a non-plateau
smooth cutoff and a nonzero open low-momentum source sector, exact source
preservation requires enlarging the Wilsonian datum to include
source-dependent vertices.  The corresponding finite-regulator object is
\begin{equation}
  \exp[-L_{\Lambda,J}^{\rm src}[\phi_<]]
  :=
  \int\dd\mu_{C_>}(\phi_>)\,
  \exp\left[-L_{\Lambda_0}[\phi_<+\phi_>]
  -\langle J,\phi_>\rangle\right],
  \label{eq:source-dependent-wilsonian-action}
\end{equation}
where \(C_{\Lambda_0}=C_\Lambda+C_>\) and
\(\langle J,\phi\rangle=\int J\phi\).  With
\[
  W_{\Lambda}^{\rm src,low}[J]
  :=
  \log
  \int\dd\mu_{C_\Lambda}(\phi_<)\,
  \exp\left[
    -L_{\Lambda,J}^{\rm src}[\phi_<]
    -\langle J,\phi_<\rangle
  \right],
\]
this gives the exact identity
\begin{equation}
  W_{\Lambda}^{\rm src,low}[J]=W_{\Lambda_0}[J]
  \label{eq:source-dependent-wilsonian-low-w-equals-bare-w}
\end{equation}
after the remaining field is integrated.  However, differentiating
\eqref{eq:source-dependent-wilsonian-low-w-equals-bare-w} with respect to
\(J\) also differentiates \(L_{\Lambda,J}^{\rm src}\).  Therefore an exact
matching theorem based on this formulation must include source-vertices as
part of the coordinate data.  The theorem below instead proves the comparison
for an \emph{admissible plateau cutoff}: choose a smooth profile satisfying
\begin{equation}
  \chi(s)=1\quad(0\le s\le c_\chi^2)
  \label{eq:admissible-plateau-cutoff}
\end{equation}
for some \(c_\chi>0\).  Let \(\mathcal E_<\) denote the finite-dimensional low
source test space chosen with
\begin{equation}
  \operatorname{supp}\widehat J\subset\{|k|<c_<\Lambda\},
  \qquad
  0<c_< < c_\chi .
  \label{eq:low-source-plateau-support}
\end{equation}
Since \(\Lambda<\Lambda_0\), this implies \(C_>(k)\widehat J(k)=0\) on the low
test space.  Thus the source-independent Wilsonian action \(L_\Lambda\) has
an exact low-source identity.  The connected functional computed from it is
denoted
\[
  W_{\Lambda}^{\rm low}[J]
  =
  \log
  \int\dd\mu_{C_\Lambda}(\phi)\,
  \exp\left[
    -L_\Lambda[\phi]
    -\int J\phi
  \right].
\]
The finite-dimensional Gaussian convolution identity and
\eqref{eq:low-source-plateau-support} give the exact equality
\begin{equation}
  W_{\Lambda}^{\rm low}[J]=W_{\Lambda_0}[J]
  \label{eq:wilsonian-low-w-equals-bare-w}
\end{equation}
at the finite regulator.  No exact source-independent statement of this form
is asserted below for non-plateau profiles such as a Gaussian cutoff.  Such a
profile may be treated either by keeping the source-dependent vertices in
\eqref{eq:source-dependent-wilsonian-action}, or by proving a separate
profile-dependent leakage estimate; that is a different comparison datum.

\begin{lemma}[Exact low-source identity for a plateau split]
\label{lem:plateau-low-source-identity}
Assume \eqref{eq:admissible-plateau-cutoff} and
\eqref{eq:low-source-plateau-support}.  At finite regulator,
\eqref{eq:wilsonian-low-w-equals-bare-w} holds for every source
\(J\in\mathcal E_<\).
\end{lemma}

\begin{proof}
By Gaussian convolution,
\[
  \int\dd\mu_{C_{\Lambda_0}}(\phi)\,
  \ee^{-L_{\Lambda_0}[\phi]-\langle J,\phi\rangle}
  =
  \int\dd\mu_{C_\Lambda}(\phi_<)
  \int\dd\mu_{C_>}(\phi_>)\,
  \ee^{-L_{\Lambda_0}[\phi_<+\phi_>]
      -\langle J,\phi_<+\phi_>\rangle}.
\]
For \(J\in\mathcal E_<\), the covariance of the scalar Gaussian random
variable \(\langle J,\phi_>\rangle\) is
\[
  \langle J,C_>J\rangle
  =
  \int_k |\widehat J(k)|^2
  \frac{\chi(k^2/\Lambda_0^2)-\chi(k^2/\Lambda^2)}{k^2+m^2}.
\]
On \(\operatorname{supp}\widehat J\), both cutoff profiles equal one by
\eqref{eq:admissible-plateau-cutoff} and
\eqref{eq:low-source-plateau-support}; hence the integrand vanishes and
\(\langle J,\phi_>\rangle=0\) almost surely.  The previous display becomes
\[
  \int\dd\mu_{C_\Lambda}(\phi_<)\,
  \ee^{-\langle J,\phi_<\rangle}
  \int\dd\mu_{C_>}(\phi_>)\,
  \ee^{-L_{\Lambda_0}[\phi_<+\phi_>]}.
\]
The inner integral is, by definition, \(\exp[-L_\Lambda[\phi_<]]\) when the
field-independent normalization is included as the vacuum-energy coordinate
of \(L_\Lambda\).  With that convention this is exactly
\(Z_\Lambda^{\rm low}[J]\), and taking logarithms gives
\eqref{eq:wilsonian-low-w-equals-bare-w}.  If one fixes the vacuum-energy
coordinate separately, the two logarithms differ only by the chosen
source-independent constant and all positive source derivatives agree.
\end{proof}

The 1PI effective action at the same finite regulator is the Legendre
transform of the connected functional.  If
\[
  \varphi(x)=\frac{\delta W[J]}{\delta J(x)},
\]
then, in the Euclidean sign convention of this volume,
\[
  \Gamma[\varphi]
  =
  W[J]-\int\dd^Dx\,J(x)\varphi(x),
  \qquad
  J=J[\varphi].
\]
Equation \eqref{eq:wilsonian-low-w-equals-bare-w} therefore implies equality
of the low-momentum 1PI kernels obtained from the source-independent
plateau-cutoff Wilsonian description and from the original regulated
description, after the same Legendre transform and the same field
normalization convention are used.  Here ``low 1PI'' means the Legendre
transform of the connected functional after restriction to the finite source
space \(\mathcal E_<\).  It is not being identified with the restriction of a
Legendre transform already taken on the full source space; that different
operation would require an additional Schur-complement statement.  The
Wilsonian vertices
\(L_\Lambda^{(n)}\) and the 1PI vertices \(\Gamma^{(n)}\) are distinct
vertices; their general relation includes all tree and loop diagrams built
from the Wilsonian vertices and the remaining cutoff covariance \(C_\Lambda\).

Figure~\ref{fig:bphz-wilsonian-one-pi-bridge} records the precise low-source
Legendre-transform bridge used here: the source-restricted connected functional
is transformed on the same finite source space, and the resulting 1PI kernels
are compared after field normalization.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      box/.style={draw,rounded corners=3pt,line width=0.8pt,
        align=center,minimum width=2.65cm,minimum height=0.85cm},
      arr/.style={-{Latex[length=2mm]},line width=0.85pt}
    }
    \node[box] (bare) at (0,1.8)
      {UV-regulated action\\at \(\Lambda_0\)};
    \node[box] (bphz) at (4.05,3.25)
      {BPHZ \(R\)-operation\\on 1PI graphs};
    \node[box] (wil) at (4.05,1.8)
      {Wilsonian action\\\(L_\Lambda\)};
    \node[box] (low) at (8.15,1.8)
      {low-source\\\(W[J]\)};
    \node[box] (leg) at (8.15,3.25)
      {Legendre transform\\\(\Gamma[\varphi]\)};
    \node[box] (coord) at (12.15,3.25)
      {1PI coordinates\\\(g_I(\mu)\)};
    \node[box] (ct) at (12.15,1.8)
      {Wilsonian coordinates\\\(u_A(\Lambda)\)};

    \draw[arr] (bare)--(bphz);
    \draw[arr] (bare)--(wil);
    \draw[arr] (wil)--(low);
    \draw[arr] (low)--(leg);
    \draw[arr] (bphz)--(leg);
    \draw[arr] (leg)--(coord);
    \draw[arr] (wil)--(ct);
    \draw[arr,dashed] (ct)--(coord);

    \node[align=center,blue!70!black] at (2.02,3.72)
      {local Taylor\\subtractions};
    \node[align=center,green!45!black] at (2.02,1.18)
      {Gaussian\\pushforward};
    \node[align=center,orange!85!black] at (10.15,1.23)
      {finite matching\\after low modes};
    \node[align=center,red!70!black] at (6.08,0.92)
      {source support\\inside plateau};
  \end{tikzpicture}
  }
  \caption{BPHZ, Wilsonian, and 1PI constructions are connected by maps.
  BPHZ acts directly on 1PI graphs, the Wilsonian action is obtained by
  Gaussian pushforward, and 1PI coordinates are obtained after the connected
  low-source functional is Legendre transformed.}
  \label{fig:bphz-wilsonian-one-pi-bridge}
\end{figure}

\begin{definition}[BPHZ--Wilsonian comparison datum]
\label{def:bphz-wilsonian-comparison-datum}
A comparison datum consists of the following finite choices.
First choose a BPHZ scheme \(\mathfrak b\): a Taylor subtraction point and
subtraction degree \(\delta(\gamma)\) for every superficially divergent
1PI subgraph type \(\gamma\), including the oversubtraction degrees used for
normal products.  The corresponding forest operation is
\[
  R_\Gamma^{\mathfrak b}I_\Gamma
  =
  \sum_{F\in\mathfrak F(\Gamma)}
  (-1)^{|F|}
  \left(\prod_{\gamma\in F} t_{\gamma}^{\delta(\gamma),\mathfrak b}\right)
  I_\Gamma ,
\]
with nested products ordered from smaller subgraphs to larger ones.  Second
choose the Wilsonian covariance split
\[
  C_{\Lambda_0}=C_\Lambda+C_{>},
\]
with an admissible plateau profile satisfying
\eqref{eq:admissible-plateau-cutoff} and a low test space satisfying
\eqref{eq:low-source-plateau-support}, the Banach chart
\(\mathcal B_{\Lambda}^{m,a}\) in which Wilsonian kernels are estimated, the
finite local coordinate projectors \(\rho_A(L_\Lambda)=u_A\), and the finite
1PI projectors \(\Pi_I\) at nonexceptional Euclidean momenta of size
\(O(\mu)\).  Third choose a matching ratio \(\zeta=\Lambda/\mu\) in a compact
subinterval of \((1,\infty)\).  All statements below are statements relative
to this datum.  A non-plateau smooth cutoff is a different datum: either
source-dependent vertices must be included, or an explicit leakage estimate
must be proved before the comparison theorem is invoked.
\end{definition}

\begin{proposition}[BPHZ local parts as Wilsonian local coordinates]
\label{prop:bphz-local-parts-wilsonian-coordinates}
Consider a power-counting finite scalar theory in a massive Euclidean
momentum-subtraction setting.  For every superficially divergent subgraph
\(\gamma\) with \(E_\gamma\) external scalar legs, the BPHZ local part
\[
  t_{\gamma}^{\delta(\gamma),\mathfrak b}\,\overline R_\gamma I_\gamma
\]
is a finite sum of external-momentum monomials
\[
  \sum_{|\alpha|\leq\delta(\gamma)}
  c_{\gamma,\alpha}^{\mathfrak b}\,p_\gamma^\alpha .
\]
After Fourier transform, each term is a local vertex of the form
\[
  c_{\gamma,\alpha}^{\mathfrak b}
  \int\dd^Dx\,
  P_{\alpha}(\partial_1,\ldots,\partial_{E_\gamma})
  \phi(x_1)\cdots\phi(x_{E_\gamma})\big|_{x_i=x}.
\]
For the minimal BPHZ degree
\(\delta(\gamma)=\omega(\gamma)\), these vertices are precisely among the
relevant or marginal local Wilsonian coordinates determined by power
counting.  Oversubtractions add a finite list of higher-derivative local
coordinates, and Zimmermann identities are the corresponding finite changes
of local operator coordinates.
\end{proposition}

\begin{proof}
The prepared integrand \(\overline R_\gamma I_\gamma\) has all proper
subdivergences subtracted.  The operation \(t_\gamma^{\delta,\mathfrak b}\)
is the Taylor polynomial in the independent external momenta of \(\gamma\) at
the specified subtraction configuration, hence it has the displayed monomial
form.  Multiplication by \(p^\alpha\) in momentum space is the Fourier
transform of a derivative acting on the fields incident to the contracted
vertex, so insertion of this Taylor polynomial into the contracted graph is
the insertion of a local interaction vertex with \(E_\gamma\) fields and
\(|\alpha|\) derivatives.

For a scalar field with engineering dimension \([\phi]=(D-2)/2\), the
dimension of such a local monomial in the derivative-free scalar examples of
this chapter is
\[
  \Delta_{\gamma,\alpha}
  =
  E_\gamma\frac{D-2}{2}+|\alpha|.
\]
The superficial degree \(\omega(\gamma)\) is precisely the largest derivative
order for which the local counterterm has scaling dimension at most \(D\), in
the power-counting convention used to define the BPHZ subtraction degree.
Thus \(|\alpha|\leq\omega(\gamma)\) is equivalent to
\(\Delta_{\gamma,\alpha}\leq D\) in this scalar setting.  These are the
relevant or marginal local coordinates in the Wilsonian linearized scaling
convention \(y_A=D-\Delta_A\).  If
\(\delta(\gamma)>\omega(\gamma)\), the additional Taylor coefficients have
\(\Delta_A>D\); they are irrelevant Wilsonian coordinates retained by the
chosen oversubtracted local chart.  The difference between two choices of
subtraction degree or finite Taylor part is therefore a finite change of
coefficients multiplying local vertices, which is the Wilsonian coordinate
form of the Zimmermann identity described in
Chapter~\ref{ch:volume-ii-subdivergences-forest-formulas}.
\end{proof}

\begin{lemma}[Low-momentum Taylor remainder from a Wilsonian norm]
\label{lem:low-momentum-taylor-remainder-wilsonian}
Fix a finite loop order \(\ell\), a finite list of low-momentum 1PI projectors
\(\Pi_I\), and a comparison interval
\[
  1<\zeta_- \leq \Lambda/\mu \leq \zeta_+ <\infty .
\]
Work in a Wilsonian Banach chart \(\mathcal B_{\Lambda}^{m,a}\) of the form
defined above, with \(m\) larger than the largest Taylor order extracted by
the projectors \(\Pi_I\).  Let \(v_{\perp,N}\) be the omitted part of
\(L_\Lambda\), and assume:
\begin{enumerate}[label=(\roman*)]
  \item \(\|v_{\perp,N}\|_{\Lambda,m,a}\leq K_N\);
  \item every Taylor monomial in \(v_{\perp,N}\) that can enter one of the
  projectors \(\Pi_I\) has excess scaling exponent at least \(p_N>0\) relative
  to the retained projection;
  \item the retained Wilsonian coordinates and the covariance
  \(C_\Lambda(k)\) stay in a bounded subset of the same chart on the
  comparison interval.
\end{enumerate}
Then, for every graph contributing through loop order \(\ell\) to one of the
projectors \(\Pi_I\) and containing at least one insertion of
\(v_{\perp,N}\), its contribution is bounded by
\[
  C_{I,\ell,N} K_N
  \left(\frac{\mu}{\Lambda}\right)^{p_N},
\]
with \(C_{I,\ell,N}\) depending on the bounded retained-coordinate set, on the
cutoff profile, on the projectors, and polynomially on
\(\log(\Lambda/\mu)\) on the comparison interval.
\end{lemma}

\begin{proof}
Consider first a single omitted vertex kernel with external momenta of size
\(O(\mu)\).  Write those momenta as \(k_i=\mu \kappa_i\), where the
\(\kappa_i\) range over a compact set fixed by the projectors.  The norm
\(\|\cdot\|_{\Lambda,m,a}\) bounds all derivatives of the dimensionless
kernel in the variables \(k_i/\Lambda\) up to order \(m\).  Since the Taylor
monomials of excess smaller than \(p_N\) have been retained and subtracted
into \(L_\Lambda^{(N)}\), Taylor's theorem with integral remainder gives
\[
  |\Pi_I v_{\perp,N}|
  \leq
  C_{I,N}K_N
  \left(\frac{\mu}{\Lambda}\right)^{p_N}
\]
for each projected vertex contribution.

For a fixed loop order \(\ell\) and a fixed finite family of projectors, only
finitely many graph topologies and Taylor derivatives are involved.  Rescale
each remaining low loop momentum by \(q=\Lambda x\).  The smooth cutoff gives
uniform bounds on the rescaled covariance and its derivatives away from the
mass singularities excluded by the massive, nonexceptional hypothesis.  The
bounded retained vertices supply constants in the chosen chart.  The
Taylor-remainder factor from at least one omitted vertex is therefore carried
through the whole graph.  Logarithms arise only from scale integrations over
the finite interval between \(\mu\) and \(\Lambda\), and on the fixed
comparison interval they are bounded by a polynomial in \(\log(\Lambda/\mu)\).
Summing the finite set of graph and derivative contributions gives the stated
constant \(C_{I,\ell,N}\).
\end{proof}

\begin{lemma}[Forest operation and the Wilsonian vertex split]
\label{lem:bphz-commutes-with-wilsonian-vertex-split}
Fix a graph \(\Gamma\) whose vertices are labeled by kernels appearing in the
additive decomposition
\[
  L_\Lambda=L_\Lambda^{(N)}+v_{\perp,N}.
\]
The BPHZ forest operation is linear in every vertex kernel and depends on the
graph only through its topology, momentum routing, subtraction degrees, and
subtraction points.  Consequently,
\[
  R_\Gamma^{\mathfrak b}I_\Gamma[L_\Lambda]
  =
  \sum_{\sigma:V(\Gamma)\to\{N,\perp\}}
  R_\Gamma^{\mathfrak b}
  I_{\Gamma,\sigma}[L_\Lambda^{(N)},v_{\perp,N}],
  \label{eq:forest-vertex-split-linearity}
\]
where \(\sigma\) records which vertices of \(\Gamma\) use retained kernels
and which use \(v_{\perp,N}\).  The summands with at least one
\(\perp\)-vertex are exactly the subtracted graphs to which
Lemma~\ref{lem:low-momentum-taylor-remainder-wilsonian} applies, provided the
lemma's Banach norm bounds are imposed on the subtracted integrands.
\end{lemma}

\begin{proof}
Before subtraction, the Feynman integrand is multilinear in the vertex
kernels.  Expanding each vertex label according to
\(L_\Lambda^{(N)}+v_{\perp,N}\) gives the finite sum over maps
\(\sigma:V(\Gamma)\to\{N,\perp\}\).  For a fixed forest \(F\), the operator
\[
  \prod_{\gamma\in F}t_\gamma^{\delta(\gamma),\mathfrak b}
\]
acts by Taylor expansion in the external variables assigned to the subgraphs
\(\gamma\).  It does not inspect whether a vertex kernel in the integrand is
retained or omitted; it is a linear differential operator with coefficients
fixed by the graph topology and by \(\mathfrak b\).  Therefore this Taylor
operator commutes with the finite additive expansion over vertex labels.
Summing over forests gives \eqref{eq:forest-vertex-split-linearity}.

If a summand contains a \(\perp\)-vertex, every forest subtraction of that
summand still contains the same \(\perp\)-vertex unless the Taylor contraction
has replaced the whole subgraph containing it by its local Taylor polynomial.
In that case the resulting local polynomial is a Taylor coefficient of a graph
with a \(v_{\perp,N}\)-insertion.  The Wilsonian Banach norm in
Lemma~\ref{lem:low-momentum-taylor-remainder-wilsonian} was imposed on the
finite set of kernels and derivatives entering these Taylor coefficients, so
the same \((\mu/\Lambda)^{p_N}\) bound applies to the subtracted summand.
\end{proof}

\begin{definition}[Admissible finite-order BPHZ--Wilsonian matching context]
\label{def:finite-order-bphz-wilsonian-matching-context}
Consider a massive Euclidean scalar theory with local polynomial
interactions, finite ultraviolet regulator \(\Lambda_0\), an intermediate
admissible plateau smooth cutoff \(\Lambda\) in the sense of
\eqref{eq:admissible-plateau-cutoff}, and nonexceptional external momenta of
size \(O(\mu)\), with \(\mu<\Lambda\).  Fix:
\begin{enumerate}[label=(\roman*)]
  \item a BPHZ subtraction scheme \(\mathfrak b\) consisting of the forest
  formula of Chapter~\ref{ch:volume-ii-subdivergences-forest-formulas} with specified
  Taylor subtraction points and oversubtraction degrees;
  \item a finite-dimensional low test space \(\mathcal E_<\) of sources and
  classical fields whose Fourier transforms are supported in
  \(|k|<c_<\Lambda\), where \(0<c_< < c_\chi\) and \(c_\chi\) is the plateau
  constant in \eqref{eq:admissible-plateau-cutoff}; this test space contains
  the momenta used by the selected 1PI projectors and satisfies \(C_>J=0\);
  \item a finite list of 1PI projectors \(\Pi_I\) at scale \(\mu\), each
  extracting a Taylor coefficient of a low-momentum 1PI kernel at a prescribed
  nonexceptional subtraction configuration \(\mathcal S_I\);
  \item a finite list of Wilsonian local-coordinate projectors \(\rho_A\) at
  scale \(\Lambda\), extracting the coefficients \(u_A=\rho_A(L_\Lambda)\) of
  local vertex tensors \(P_A\).
\end{enumerate}
Let \(L_\Lambda^{(N)}(u)\) be the local Wilsonian action reconstructed from
the retained coordinates \(u_A\), and let
\[
  v_{\perp,N}:=L_\Lambda-L_\Lambda^{(N)}(u)
\]
be the omitted part.  For a fixed loop order \(\ell\), define the matching
map
\begin{equation}
  M_I^{(\ell,N)}(u;\Lambda,\mu,\mathfrak b)
  :=
  \Pi_I\,
  \Gamma^{<,(\leq\ell)}_{\Lambda,N}[u],
  \label{eq:finite-order-bphz-wilsonian-matching-map}
\end{equation}
where \(\Gamma^{<,(\leq\ell)}_{\Lambda,N}[u]\) is the low-mode 1PI effective
action defined as follows: compute the connected functional by integrating
the full finite-regulator remaining field with covariance \(C_\Lambda\) and
vertices \(L_\Lambda^{(N)}(u)\), restrict its source argument to
\(\mathcal E_<\), then take the local Legendre transform on that finite
source space with the same finite local normalization convention as
\(\mathfrak b\).  The resulting formal 1PI functional is truncated after loop
order \(\ell\).
\end{definition}

\begin{hypothesis}[Finite-order BPHZ--Wilsonian matching estimates]
\label{hyp:finite-order-bphz-wilsonian-matching-estimates}
In the context of
Definition~\ref{def:finite-order-bphz-wilsonian-matching-context}, assume the
following estimates hold on the coordinate neighborhood under consideration:
\begin{enumerate}[label=(\alph*)]
  \item the BPHZ forest theorem applies to all graphs through loop order
  \(\ell\) that contribute to the projectors \(\Pi_I\);
  \item the Hessian of \(W_\Lambda^{\rm low}\) on the low source space is
  invertible on the source neighborhood used for the Legendre transform;
  \item in a chosen Wilsonian Banach norm,
  \(\|v_{\perp,N}\|_{\Lambda,m,a}\leq K_N\), and the first omitted local
  tensor that can contribute to the projectors \(\Pi_I\) has excess scaling
  exponent at least \(p_N>0\), with the retained coordinates and covariance
  bounded as in Lemma~\ref{lem:low-momentum-taylor-remainder-wilsonian};
  the same bounds hold for the finite set of Taylor-subtracted kernels
  obtained by applying \(R_\Gamma^{\mathfrak b}\) to graphs with at least one
  \(v_{\perp,N}\)-insertion.
  This bounded-chart hypothesis is an order-by-order perturbative hypothesis
  in the renormalized coupling, or a genuine nonperturbative hypothesis only
  for trajectories for which the stated norm remains bounded;
  \item the limit \(\Lambda_0\to\infty\), when it is taken, is taken along a
  Polchinski-flow counterterm trajectory: the relevant and marginal
  UV-cutoff coordinates in the regulated path-integral Lagrangian at
  \(\Lambda_0\) are tuned as functions of \(\Lambda_0\) so that the selected
  renormalized boundary coordinates at a fixed reference scale stay fixed and
  the Wilsonian kernels at the intermediate scale \(\Lambda\) have the
  perturbative limits supplied by Polchinski/Salmhofer/Kopper--M\"uller-type
  cutoff estimates in the chosen norm.
\end{enumerate}
\end{hypothesis}

\begin{theorem}[Finite-order BPHZ--Wilsonian matching]
\label{thm:finite-order-bphz-wilsonian-matching}
In an admissible matching context as in
Definition~\ref{def:finite-order-bphz-wilsonian-matching-context}, and under
Hypothesis~\ref{hyp:finite-order-bphz-wilsonian-matching-estimates}, the
BPHZ-renormalized 1PI coordinates through loop order \(\ell\) obey
\begin{equation}
  g_I^{\mathfrak b,(\leq\ell)}(\mu)
  =
  M_I^{(\ell,N)}(u(\Lambda);\Lambda,\mu,\mathfrak b)
  +
  R_I^{(\ell,N)}(\Lambda,\mu),
  \qquad
  |R_I^{(\ell,N)}|
  \leq
  C_{I,\ell,N}K_N
  \left(\frac{\mu}{\Lambda}\right)^{p_N}.
  \label{eq:finite-order-bphz-wilsonian-coordinate-comparison}
\end{equation}
If a sublist \(A_0\) of Wilsonian coordinates has the same cardinality as the
chosen 1PI coordinates and the Jacobian
\[
  J_{IA}
  =
  \frac{\partial M_I^{(\ell,N)}}{\partial u_A}
  \bigg|_{u=u_*},
  \qquad
  I\in\mathcal I,\ A\in A_0,
\]
is invertible at a reference point \(u_*\), then the finite-dimensional
implicit function theorem gives a local finite change of coordinates
\[
  u_A
  =
  G_A\bigl(g^{\mathfrak b},u_{\bar A};\Lambda/\mu,\mathfrak b\bigr),
  \qquad A\in A_0,
\]
with the same \(O((\mu/\Lambda)^{p_N})\) truncation error.
\end{theorem}

\begin{proof}
Lemma~\ref{lem:plateau-low-source-identity} gives
\eqref{eq:wilsonian-low-w-equals-bare-w} for all sources in \(\mathcal E_<\).
This equality is an equality of functions on the same finite-dimensional
source space.  By
Hypothesis~\ref{hyp:finite-order-bphz-wilsonian-matching-estimates}\textup{(b)},
the maps
\[
  J\longmapsto \varphi=\frac{\delta W[J]}{\delta J}
\]
are locally invertible for both sides of
\eqref{eq:wilsonian-low-w-equals-bare-w}.  Since the functions \(W\) agree,
their first derivatives and local inverse maps agree.  Therefore the
Legendre transforms on \(\mathcal E_<\),
\[
  \Gamma^<[\varphi]=W[J(\varphi)]-\langle J(\varphi),\varphi\rangle ,
\]
agree on the corresponding low classical-field neighborhood.  This proves
equality of low-mode 1PI kernels at finite regulator.  The object just
constructed is the 1PI functional for the low test space; it is the object
used by the projectors \(\Pi_I\).

Theorem~\ref{thm:bphz-finiteness-massive-euclidean-scalar} proves, by the
Hepp-sector Taylor-remainder estimates in
Chapter~\ref{ch:volume-ii-subdivergences-forest-formulas}, that at the fixed
loop order \(\ell\) every ultraviolet subtraction in the chosen massive BPHZ
scheme \(\mathfrak b\) is a Taylor polynomial in the external variables of
the subgraph and that the subtracted graph has a finite
\(\Lambda_0\to\infty\) limit.  Thus, after the selected subtraction points,
field normalization, and oversubtraction degrees are fixed, the projected
BPHZ coordinates \(g_I^{\mathfrak b,(\leq\ell)}(\mu)\) are finite Taylor
coefficients of the same low-mode 1PI kernels obtained from the Wilsonian
action.  The \(\Lambda_0\to\infty\) limit here is not taken at fixed
UV-cutoff Lagrangian coordinates: by
Hypothesis~\ref{hyp:finite-order-bphz-wilsonian-matching-estimates}\textup{(d)},
it is taken along the
Polchinski-flow counterterm trajectory that holds the chosen renormalized
boundary data fixed and has a limit for the Wilsonian kernels at the
intermediate scale \(\Lambda\).  At fixed loop order this is precisely the
perturbative renormalized trajectory constructed by the Wilsonian estimates,
not a nonperturbative assertion about a cutoff coupling held fixed.

Computing the projected low kernels with the retained local action
\(L_\Lambda^{(N)}(u)\) is exactly the definition
\eqref{eq:finite-order-bphz-wilsonian-matching-map} of
\(M_I^{(\ell,N)}\).

It remains only to compare the exact Wilsonian action with its finite local
projection.  The graph expansion is first decomposed according to
\(L_\Lambda=L_\Lambda^{(N)}+v_{\perp,N}\).  By
Lemma~\ref{lem:bphz-commutes-with-wilsonian-vertex-split}, the BPHZ forest
operation commutes with this finite additive vertex decomposition, and the
subtracted graphs with at least one \(v_{\perp,N}\)-insertion are still graphs
to which the Wilsonian Taylor-remainder norm estimate applies.
Hypothesis~\ref{hyp:finite-order-bphz-wilsonian-matching-estimates}\textup{(c)}
puts those subtracted graphs in the situation of
Lemma~\ref{lem:low-momentum-taylor-remainder-wilsonian}.  The lemma gives the
displayed power bound for their contribution to the selected Taylor
coefficients of low 1PI kernels.  Substituting this estimate into the equality
of finite-regulator low 1PI kernels, and then taking the tuned
BPHZ/Polchinski \(\Lambda_0\to\infty\) limit at fixed \(\Lambda\) and
\(\mu\), gives
\eqref{eq:finite-order-bphz-wilsonian-coordinate-comparison}.

The final assertion is finite-dimensional.  If \(J_{IA}\) is invertible, the
map from the selected Wilsonian coordinates \(u_A\), \(A\in A_0\), to the
selected 1PI coordinates has an invertible derivative at \(u_*\).  The
implicit function theorem therefore gives the displayed local coordinate
map \(G_A\), with the same remainder estimate because the previous paragraph
already controls the omitted Wilsonian coordinates.
\end{proof}

The finite change of coordinates in
Theorem~\ref{thm:finite-order-bphz-wilsonian-matching} can be constructed
recursively in the loop expansion.  Write
\[
  M_I^{(\ell,N)}(u)
  =
  \sum_{r=0}^{\ell}\hbar^r M_{I,r}(u),
  \qquad
  u_A
  =
  \sum_{r=0}^{\ell}\hbar^r U_{A,r},
  \qquad
  g_I
  =
  \sum_{r=0}^{\ell}\hbar^r g_{I,r}.
\]
Choose the sublist \(A_0\) with invertible tree-level Jacobian
\[
  J_{IA}=
  \frac{\partial M_{I,0}}{\partial u_A}\bigg|_{u=U_0},
  \qquad A\in A_0 .
\]
At order zero, solve \(g_{I,0}=M_{I,0}(U_0)\).  Suppose that
\(U_{A,s}\) has been determined for \(s<r\).  The coefficient of
\(\hbar^r\) in \(M_I(u)\) has the form
\[
  [\hbar^r]\,M_I(u)
  =
  J_{IA}U_{A,r}
  +
  \mathcal F_{I,r}(U_0,\ldots,U_{r-1};u_{\bar A}),
\]
where \(\mathcal F_{I,r}\) is computed from lower-order coordinate
coefficients, the explicit \(r\)-loop low-mode 1PI graphs, and the retained
coordinates \(u_{\bar A}\) outside \(A_0\).  Therefore
\[
  U_{A,r}
  =
  (J^{-1})_{AI}
  \left[
    g_{I,r}
    -
    \mathcal F_{I,r}(U_0,\ldots,U_{r-1};u_{\bar A})
  \right].
\]
This recursion is the perturbative coordinate map
\[
  u_A=G_A(g,u_{\bar A};\Lambda/\mu,\mathfrak b)
\]
order by order.  The coefficients
\(\mathcal F_{I,r}\) are finite because BPHZ supplies finite 1PI projectors at
the chosen loop order, and the Wilsonian Taylor-remainder lemma controls the
omitted local coordinates by the displayed power of \(\mu/\Lambda\).

\begin{remark}[Scope of the matching theorem]
The theorem is a fixed-loop perturbative statement in a specified subtraction
scheme, a specified Wilsonian norm, and the admissible plateau cutoff
\eqref{eq:admissible-plateau-cutoff}.  The plateau is load-bearing: it makes
\(C_>J=0\) on the low test space and hence makes
\eqref{eq:wilsonian-low-w-equals-bare-w} exact for the source-independent
Wilsonian action used in the matching map.  Non-plateau smooth profiles such
as Gaussian cutoffs require either the source-dependent Wilsonian action
\eqref{eq:source-dependent-wilsonian-action} together with source-vertex
coordinates, or a separate leakage theorem; they are not covered by
Theorem~\ref{thm:finite-order-bphz-wilsonian-matching} as stated.

The massive BPHZ scheme used here is
the forest-formula scheme of
Chapter~\ref{ch:volume-ii-subdivergences-forest-formulas}.  Massless theories require
BPHZL or another infrared-safe subtraction scheme, and dimensional
regularization with minimal subtraction is a different coordinate system whose
relation to the present one is another finite matching map.  Since BPHZ is a
perturbative forest-formula construction, there is no separate
``nonperturbative BPHZ construction'' being compared here.  The
nonperturbative question is instead whether a Wilsonian or constructive
continuum theory exists and whether its correlation functions have, in the
chosen local coordinates, the BPHZ-renormalized series as their perturbative
asymptotic expansion.  That question requires existence of the relevant
measures or functionals, Legendre transforms, and cutoff-uniform Banach
estimates for the model under study.

The bounded-chart hypothesis in the theorem should also not be read as a
claim about arbitrary bare trajectories.  For non-asymptotically-free models
such as four-dimensional scalar quartic theory, the nonperturbative continuum
limit at fixed positive bare family is obstructed in the standard
reflection-positive lattice universality classes discussed earlier.  The
Polchinski theorem used here is an order-by-order theorem in the renormalized
coupling after the counterterms have been tuned; in that perturbative
coefficientwise sense the retained coordinates remain in the required
neighborhood at every fixed loop order.
\end{remark}

\begin{remark}[Status of the comparison]
The exact finite-regulator input is
\eqref{eq:wilsonian-low-w-equals-bare-w}, valid here because the cutoff has a
low-momentum plateau and the source support is strictly inside it.  The
Legendre-transform step is a finite-dimensional local theorem under Hessian
invertibility on the chosen low test space.  The BPHZ input is the
order-by-order massive forest-formula theorem proved in
Chapter~\ref{ch:volume-ii-subdivergences-forest-formulas}, with BPHZL or
another infrared prescription required for massless exceptional regions.  The
Wilsonian input is an estimate in the chosen normed RG chart and along the
tuned Polchinski-flow counterterm trajectory.  Minimal subtraction, momentum
subtraction, and Wilsonian coordinates are different coordinate systems; their
comparison is the finite matching problem just constructed.  A theorem
connecting this perturbative comparison to an actual continuum QFT requires
additional existence and uniform-estimate theorems for the model under study.
\end{remark}

\begin{corollary}[Fixed-loop comparison of renormalizability criteria]
\label{cor:fixed-loop-bphz-polchinski-renormalizability}
In the massive, nonexceptional, finite-projector, plateau-cutoff, and tuned
Polchinski-trajectory setting of
Theorem~\ref{thm:finite-order-bphz-wilsonian-matching}, the following two
renormalizability criteria are equivalent through the fixed loop order
\(\ell\), up to the displayed
\((\mu/\Lambda)^{p_N}\) error from omitted irrelevant Wilsonian coordinates.
\begin{enumerate}[label=(\roman*)]
  \item BPHZ renormalizability: after applying the forest formula in the
  chosen scheme \(\mathfrak b\), all projected 1PI coordinates
  \(g_I^{\mathfrak b,(\leq\ell)}(\mu)\) have finite
  \(\Lambda_0\to\infty\) limits, and the required subtraction freedom is a
  finite list of local Taylor coefficients.
  \item Polchinski-Wilsonian renormalizability: the Wilsonian trajectory is
  specified at scale \(\Lambda\) by a finite list of relevant or marginal local
  coordinates \(u_A\), together with an irrelevant remainder whose norm obeys
  Lemma~\ref{lem:low-momentum-taylor-remainder-wilsonian}; the selected low
  1PI coordinates are then finite functions of \(u_A\) after low-mode
  integration.
\end{enumerate}
First, BPHZ subtraction data determine Wilsonian local coordinates by
Proposition~\ref{prop:bphz-local-parts-wilsonian-coordinates}.  The selected
1PI coordinates are then obtained by the finite matching map
\(M_I^{(\ell,N)}\).  Conversely, a Wilsonian datum satisfying the stated norm
bound determines finite BPHZ projected coordinates: apply the same projectors
\(\Pi_I\) to the low-mode 1PI functional, and assign the resulting local Taylor
parts to the BPHZ subtraction coordinates.  A change of BPHZ subtraction point,
oversubtraction degree, or finite Taylor part changes the Wilsonian description
by a finite local coordinate transformation in the retained chart.
\end{corollary}

\begin{proof}
Assume first the BPHZ criterion.  Proposition
\ref{prop:bphz-local-parts-wilsonian-coordinates} identifies every forest
Taylor subtraction through the selected loop order with a local vertex in the
Wilsonian chart.  Minimal subtractions occupy relevant and marginal
coordinates, while oversubtractions occupy the explicitly retained
higher-derivative local coordinates.  After these local coordinates are fixed,
Theorem~\ref{thm:finite-order-bphz-wilsonian-matching} gives
\[
  g_I^{\mathfrak b,(\leq\ell)}(\mu)
  =
  M_I^{(\ell,N)}(u(\Lambda);\Lambda,\mu,\mathfrak b)
  +
  R_I^{(\ell,N)}
\]
with the stated bound on \(R_I^{(\ell,N)}\).  Thus the BPHZ finite local
subtraction freedom is represented by the finite Wilsonian coordinate list,
and the omitted Wilsonian coordinates affect the selected projectors only by
the Taylor-remainder estimate.

Conversely, assume the Wilsonian criterion.  The finite-regulator Gaussian
identity constructs the same low-source connected functional as the original
regulated theory, and the Hessian hypothesis in
Theorem~\ref{thm:finite-order-bphz-wilsonian-matching} constructs the low-mode
Legendre transform.  Applying the projectors \(\Pi_I\) to this low-mode 1PI
functional gives finite numbers through loop order \(\ell\).  For every
superficially divergent graph contributing to these projectors, its local
Taylor part is one of the finitely many local monomials in the retained
Wilsonian chart, again by Proposition
\ref{prop:bphz-local-parts-wilsonian-coordinates}.  Assigning those Taylor
coefficients as BPHZ subtraction coordinates gives a BPHZ scheme whose
projected coordinates agree with the Wilsonian values up to the same
remainder estimate.  The construction is recursive in loop order because the
forest formula subtracts proper divergent subgraphs before the overall graph,
whereas the Wilsonian matching recursion displayed above solves the local
coordinate map at order \(r\) using only lower-order coordinate data and the
explicit \(r\)-loop low-mode graphs.

Finally, the difference between two BPHZ subtraction choices is a finite sum
of Taylor polynomials in external momenta.  Fourier transforming these
polynomials gives a finite sum of local vertices, hence a finite change of
the retained Wilsonian coordinates.  Where the Jacobian in
Theorem~\ref{thm:finite-order-bphz-wilsonian-matching} is invertible, this
finite change is a local coordinate transformation by the implicit function
theorem.
\end{proof}

\begin{remark}[What is, and is not, proved]
Corollary~\ref{cor:fixed-loop-bphz-polchinski-renormalizability} proves a
two-way implication only after the finite-loop, massive, nonexceptional, and
finite-projector hypotheses, together with the plateau-cutoff and tuned
counterterm-trajectory hypotheses, of
Theorem~\ref{thm:finite-order-bphz-wilsonian-matching} have been imposed.  In
that setting, BPHZ forest-formula local Taylor parts construct the Wilsonian
local coordinates, and controlled Wilsonian local coordinates construct BPHZ
subtraction coordinates by projecting the low-mode 1PI functional.  This is
an equivalence of finite coordinate descriptions of selected perturbative
1PI data, with the omitted irrelevant coordinates contributing the stated
\((\mu/\Lambda)^{p_N}\) remainder.

The corollary is not a theorem that a nonperturbatively constructed
Wilsonian or constructive continuum theory exists, nor that such a theory has
the above perturbative series as its asymptotic expansion.  BPHZ itself is
already a perturbative construction.  The additional nonperturbative problem
would require, for the model in question, a construction of the regulated and
limiting functionals, a topology on the full Wilsonian coordinate space,
cutoff-uniform estimates in that topology, and a well-defined Legendre
transform on the required source domain.  The fixed-loop result is the part
that can be proved from the perturbative forest formula and the Wilsonian
cutoff estimates displayed in this chapter.
\end{remark}

\section{One-Loop Matching in the Quartic Scalar Theory}

The theorem's matching map can be displayed directly in the four-point
function of the massive scalar theory in four Euclidean dimensions.  Work with
\[
  S_{\rm int}[\phi]
  =
  \frac{g_0}{4!}\int\dd^4x\,\phi(x)^4
\]
and a finite covariance \(C_{\Lambda_0}(k)\).  Let
\[
  C_{\Lambda_0}(k)=C_\Lambda(k)+C_{>} (k),
  \qquad
  C_{>}(k)=C_{\Lambda_0}(k)-C_\Lambda(k),
\]
where \(C_\Lambda\) is the covariance of the remaining low field and
\(C_>\) is the covariance integrated out in the Wilsonian step from
\(\Lambda_0\) to \(\Lambda\).  For a four-point external configuration
\[
  \mathcal P=(p_1,p_2,p_3,p_4),
  \qquad
  \sum_{i=1}^4p_i=0,
\]
define the channel momenta
\[
  P_s=p_1+p_2,\qquad
  P_t=p_1+p_3,\qquad
  P_u=p_1+p_4 .
\]
For any two covariances \(A,B\), define the one-loop bubble functional
\[
  \mathcal B_{A,B}(\mathcal P)
  =
  \sum_{c=s,t,u}
  \int\frac{\dd^4q}{(2\pi)^4}\,
  A(q)\,B(q+P_c).
\]
Thus \(\mathcal B_{C_{\Lambda_0},C_{\Lambda_0}}\) is the regulated full
one-loop bubble sum, while
\[
  \mathcal B_{<,<}:=\mathcal B_{C_\Lambda,C_\Lambda},
  \qquad
  \mathcal B_{>,>}:=\mathcal B_{C_>,C_>}
\]
record the low-low and high-high parts of the loop.  The remaining mixed
piece is
\[
  \mathcal B_\times
  =
  \mathcal B_{C_\Lambda,C_>}
  +
  \mathcal B_{C_>,C_\Lambda}.
\]
The exact finite-regulator identity is
\begin{equation}
  \mathcal B_{C_{\Lambda_0},C_{\Lambda_0}}
  =
  \mathcal B_{<,<}
  +
  \mathcal B_{>,>}
  +
  \mathcal B_\times .
  \label{eq:bubble-covariance-split}
\end{equation}

In the Wilsonian action, the high-high part appears directly in the quartic
vertex.  Let \(u_4(\Lambda;\mathcal P)\) be the coefficient whose contribution
to \(L_\Lambda\) has four-field kernel \(u_4\), so that the corresponding
tree 1PI vertex is \(-u_4\).  To second order in \(g_0\),
\begin{equation}
  u_4(\Lambda;\mathcal P)
  =
  g_0
  -
  \frac{g_0^2}{2}\,
  \mathcal B_{>,>}(\mathcal P)
  +
  u^{\rm loc}_{4,{\rm ct}}(\Lambda;\mathcal P)
  +
  O(g_0^3),
  \label{eq:wilsonian-quartic-one-loop-high-bubble}
\end{equation}
where \(u^{\rm loc}_{4,{\rm ct}}\) denotes the local quartic counterterm and
finite local convention chosen at the ultraviolet regulator.  The factor
\(1/2\) is the symmetry factor of the bubble in the normalization
\((g_0/4!)\phi^4\).

Figure~\ref{fig:quartic-one-loop-bphz-wilsonian-matching} displays the one-loop
matching calculation: the high-momentum bubble in the Wilsonian quartic vertex
is compared with the BPHZ-subtracted 1PI quartic coordinate after the chosen
finite local convention is fixed.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      arr/.style={-{Latex[length=2mm]},line width=0.85pt},
      box/.style={draw,rounded corners=3pt,align=center,
        minimum width=2.65cm,minimum height=0.76cm}
    }
    \node[box] (full) at (0,1.8)
      {full regulated\\bubble};
    \node[box] (hh) at (4.0,3.0)
      {high--high\\part};
    \node[box] (ll) at (4.0,1.8)
      {low--low\\part};
    \node[box] (mix) at (4.0,0.6)
      {mixed\\part};
    \node[box] (wil) at (8.1,3.0)
      {Wilsonian\\quartic vertex};
    \node[box] (low) at (8.1,1.8)
      {low-field\\1PI loop};
    \node[box] (loc) at (8.1,0.6)
      {local matching\\coordinates};
    \node[box] (coord) at (12.0,1.8)
      {1PI coordinate\\at \(\mathcal S_\mu\)};

    \draw[arr] (full)--(hh);
    \draw[arr] (full)--(ll);
    \draw[arr] (full)--(mix);
    \draw[arr] (hh)--(wil);
    \draw[arr] (ll)--(low);
    \draw[arr] (mix)--(loc);
    \draw[arr] (wil)--(coord);
    \draw[arr] (low)--(coord);
    \draw[arr] (loc)--(coord);

    \node[align=center,blue!70!black] at (2.0,3.42)
      {\(C_>C_>\)};
    \node[align=center,green!45!black] at (2.0,2.18)
      {\(C_\Lambda C_\Lambda\)};
    \node[align=center,orange!85!black] at (1.75,0.55)
      {\(C_\Lambda C_>\)\\\(+C_>C_\Lambda\)};
  \end{tikzpicture}
  }
  \caption{At one loop the regulated four-point bubble splits according to the
  covariance decomposition.  The high--high part contributes to the Wilsonian
  quartic vertex, the low--low part is produced by the remaining low-field
  Legendre transform, and the mixed part is represented by local matching data
  in the low-momentum expansion.}
  \label{fig:quartic-one-loop-bphz-wilsonian-matching}
\end{figure}

The mixed term in \eqref{eq:bubble-covariance-split} must be accounted for
explicitly.  In the exact Wilsonian functional it is not absent: it is
produced by lower Wilsonian vertices generated by one high contraction and
then used in the remaining low-field integration.  If the external momenta are
\(O(\mu)\) with \(\mu\ll\Lambda\), this contribution is smooth in the
external invariants and has a Taylor expansion in powers of \(\mu/\Lambda\).
In a finite derivative projection it is therefore part of the local matching
map and of the first omitted irrelevant coordinates.

Let \(\mathcal S_\mu\) be the symmetric Euclidean subtraction configuration
used to define the 1PI quartic coordinate
\[
  g(\mu)=-\Gamma^{(4)}(\mathcal S_\mu).
\]
For the logarithmically divergent one-loop four-point graph, the BPHZ overall
subtraction is subtraction of the zeroth Taylor coefficient at
\(\mathcal S_\mu\).  Thus, after the coupling counterterm has been fixed by
the displayed condition,
\begin{equation}
  \Gamma^{(4)}_{\rm BPHZ}(\mathcal P)
  =
  -g(\mu)
  +
  \frac{g(\mu)^2}{2}
  \left[
    \mathcal B_{C_{\Lambda_0},C_{\Lambda_0}}(\mathcal P)
    -
    \mathcal B_{C_{\Lambda_0},C_{\Lambda_0}}(\mathcal S_\mu)
  \right]
  +
  O(g^3),
  \label{eq:bphz-quartic-one-loop-subtracted}
\end{equation}
with the limit \(\Lambda_0\to\infty\) taken after the subtraction.  The
difference in brackets has a finite limit at nonexceptional momenta.

The same finite-regulator 1PI kernel computed from the Wilsonian action is
\begin{equation}
\begin{aligned}
  \Gamma^{(4)}_{\Lambda,{\rm low}}(\mathcal P)
  &=
  -u_4(\Lambda;\mathcal P)
  +
  \frac{u_4(\Lambda;\mathcal S_\mu)^2}{2}
  \mathcal B_{<,<}(\mathcal P)        \\
  &\quad
  +
  \mathcal M_\times(\Lambda;\mathcal P)
  +
  O(u_4^3).
\end{aligned}
  \label{eq:wilsonian-low-quartic-one-loop-kernel}
\end{equation}
Here \(\mathcal M_\times\) is the contribution generated by the mixed
covariance assignments, expressed in the low-field description.  Its Taylor
polynomial at \(\mathcal S_\mu\) is local.  Projecting
\eqref{eq:wilsonian-low-quartic-one-loop-kernel} at \(\mathcal S_\mu\) gives
the one-loop matching map
\begin{equation}
  g(\mu)
  =
  u_4(\Lambda;\mathcal S_\mu)
  -
  \frac{u_4(\Lambda;\mathcal S_\mu)^2}{2}
  \mathcal B_{<,<}(\mathcal S_\mu)
  -
  \Pi_{\mathcal S_\mu}\mathcal M_\times(\Lambda)
  +
  O(u_4^3),
  \label{eq:one-loop-quartic-wilsonian-to-one-pi-coordinate}
\end{equation}
where \(\Pi_{\mathcal S_\mu}\) denotes evaluation of the four-point 1PI
projector defining the coupling.  If the Wilsonian derivative expansion is
truncated after the local quartic coordinate, the finite matching map
\eqref{eq:finite-order-bphz-wilsonian-matching-map} shifts the retained
coordinate by
\[
  -\Pi_{\mathcal S_\mu}\mathcal M_\times(\Lambda)+O(u_4^3).
\]
If the first discarded Taylor coefficient in the four-point kernel has excess
exponent \(p_N\), the Taylor-remainder estimate of
Lemma~\ref{lem:low-momentum-taylor-remainder-wilsonian} gives the bound
\[
  C_N\|v_{\perp,N}\|_{\Lambda,m,a}
  \left(\frac{\mu}{\Lambda}\right)^{p_N}.
\]

Subtracting the value at \(\mathcal S_\mu\) from
\eqref{eq:wilsonian-low-quartic-one-loop-kernel} and using the matching map
removes the local coordinate ambiguity.  The nonlocal low-momentum dependence
is
\[
  \Gamma^{(4)}_{\Lambda,{\rm low}}(\mathcal P)
  -
  \Gamma^{(4)}_{\Lambda,{\rm low}}(\mathcal S_\mu)
  =
  \frac{g(\mu)^2}{2}
  \left[
    \mathcal B_{<,<}(\mathcal P)
    -
    \mathcal B_{<,<}(\mathcal S_\mu)
  \right]
  +
  \hbox{local Taylor terms}
  +
  O(g^3).
\]
Adding the local high and mixed Taylor data reconstructs the low-momentum
expansion of the BPHZ expression
\eqref{eq:bphz-quartic-one-loop-subtracted}.  This calculation displays the
three roles separately: BPHZ removes the ultraviolet part of the full bubble
by a local subtraction; the Wilsonian vertex stores the high-momentum part of
the same graph; the 1PI coordinate is obtained only after the remaining
low-field Legendre transform and the projector at \(\mathcal S_\mu\).

The comparison proposition and the one-loop calculation also fix the role of
beta functions.  The BPHZ construction
proves locality and finiteness of the subtracted 1PI kernels.  The 1PI
renormalization group differentiates finite coordinates \(g_I(\mu)\) with
respect to the subtraction scale \(\mu\).  The Wilsonian flow differentiates
the cutoff-dependent action \(L_\Lambda\) with respect to \(\Lambda\).  These
two vector fields live on different coordinate spaces.  They become comparable
only after the matching map from \(L_\Lambda\) to the 1PI coordinates
\(g_I(\mu)\) has been specified.  Under a finite change of coordinates,
their components transform by the chain rule, as in the 1PI scheme-change
discussion.  Equality of physical low-momentum correlators is therefore a
weaker statement than equality of the displayed beta-function components in
two unmatched coordinate systems.

For gauge theories the same diagram of constructions must be lifted to the
BRST/BV setting.  The Wilsonian pushforward must preserve the quantum master
equation, and the 1PI Legendre transform must satisfy the corresponding
Slavnov--Taylor or BV master identity.  That gauge-theoretic refinement is
developed in the BV chapter; the present scalar discussion supplies the
renormalization-theoretic comparison without adding gauge constraints.
