\chapter{The Standard Model as a Hybrid Quantum Field Theory}
\label{ch:standard-model-hybrid}
\ConventionsLorentzian

The Standard Model used in physics is not presently a single completed
constructive definition of a four-dimensional continuum QFT.  It is a precise
hybrid object whose components have different mathematical statuses:
nonperturbative QCD, understood through its continuum Wightman target and
through lattice regulators used to define hadronic matrix elements; the
renormalized electroweak chiral gauge theory, defined order by order in
perturbation theory in the same sense in which perturbative QED is defined;
hybrid coupling data that pairs perturbative electroweak fields and sources
with specified gauge-invariant local QCD operators; lower-energy weak
effective theories; and matching maps between overlapping regimes.  No
physical ultraviolet cutoff is a parameter of a Standard Model prediction.
Cutoffs and regulators enter either as auxiliary devices that are removed or
absorbed into renormalized coordinates, or as extra data in a different
Wilsonian or SMEFT problem.  The purpose of this chapter is to write the
hybrid object without hiding these domains of validity.  Within each domain
the standard formulae are derived as statements about explicit fields,
couplings, operators, and observables.  Assertions about higher energies, large
field values, or physics beyond this field content require additional
mathematical data; they are not supplied merely by drawing a perturbative RG
curve or by writing a formal Lagrangian.

The local representation theory is developed first, because it fixes the
charges, anomalies, mass coordinates, and operator coordinates that every
perturbative, effective, or nonperturbative realization must reproduce.  The
hybrid definition of the full object is then assembled after these components
have been made explicit.  Until that point, the phrase ``the Standard Model''
in a local formula means the corresponding renormalized perturbative
electroweak construction together with the stated QCD and matching data, not an
assumed constructive theorem and not a finite-cutoff theory.

\section{Gauge and Matter Data}

\begin{convention}[Electroweak charge and trace normalizations]
\label{conv:sm-electroweak-trace-normalization}
The hypercharge coordinate is normalized by
\[
  Q_{\rm em}=T^3_L+Y ,
\]
where \(T^3_L\) is the third generator of the weak \(SU(2)_L\) doublet in the
standard electroweak normalization.  For transparent mass and charge formulae,
this chapter temporarily uses the conventional explicit-coupling fields
\(\mathcal G_\mu,\mathcal W_\mu,\mathcal B_\mu\).  The conversion to the
monograph's Yang--Mills convention of
Chapter~\ref{ch:vol2-classical-yang-mills-matter} is
\[
  A^{(3)}_\mu=g_3\mathcal G^A_\mu T^A,\qquad
  A^{(2)}_\mu=g_2\mathcal W^a_\mu \tau^a,\qquad
  A^{(1)}_\mu=g_1\mathcal B_\mu Y ,
\]
where \(T^A\) and \(\tau^a=\sigma^a/2\) are the conventional half-trace
generators, normalized by
\(\operatorname{tr}(\tau^a\tau^b)=\frac12\delta^{ab}\) in the fundamental
weak doublet.  If \(t^a=\sqrt2\,\tau^a\) denotes the trace-delta basis used in
Chapter~\ref{ch:vol2-classical-yang-mills-matter}, then
\(g_{2,\rm mono}=g_2/\sqrt2\) in a matter vertex.  The invariant kinetic
terms are equivalently written as explicit-coupling kinetic terms for
\(\mathcal G,\mathcal W,\mathcal B\) or as
\(-\frac1{4g^2}\operatorname{tr}(F_{\mu\nu}F^{\mu\nu})\) for the
coupling-absorbed connections.
\end{convention}

\begin{definition}[One-generation left-Weyl representation]
\label{def:sm-one-generation-left-weyl}
For one generation, written as left-handed Weyl fields, the representation is
\[
\begin{array}{c|c|c}
  \text{field} & SU(3)_c\times SU(2)_L & Y \\ \hline
  Q_L=(u_L,d_L) & (\mathbf 3,\mathbf 2) & 1/6\\
  u^c & (\overline{\mathbf 3},\mathbf 1) & -2/3\\
  d^c & (\overline{\mathbf 3},\mathbf 1) & 1/3\\
  L_L=(\nu_L,e_L) & (\mathbf 1,\mathbf 2) & -1/2\\
  e^c & (\mathbf 1,\mathbf 1) & 1 .
\end{array}
\]
Here \(u^c,d^c,e^c\) are left-handed charge-conjugate fields, not
right-handed fields in the same representation.  A gauge-singlet neutrino
field, if included in the same left-handed convention as \(\nu^c\), has
representation \((\mathbf1,\mathbf1)_0\) and does not affect the local gauge
anomaly sums.
\end{definition}

\begin{definition}[Minimal Higgs and Yukawa sector]
\label{def:sm-higgs-yukawa-sector}
The minimal Higgs scalar is
\[
  H\in(\mathbf1,\mathbf2)_{1/2}.
\]
Set
\[
  \widetilde H:=\ii\sigma^2H^*,
  \qquad
  \widetilde H\in(\mathbf1,\mathbf2)_{-1/2},
\]
where the second statement uses the pseudoreality of the \(SU(2)\) doublet.
In the left-handed Weyl convention of
Definition~\ref{def:sm-one-generation-left-weyl}, the renormalizable Yukawa
sector is
\[
  \mathcal L_Y
  =
  - (Y_u)_{IJ}\,(Q_L^I H)\,u^{cJ}
  - (Y_d)_{IJ}\,(Q_L^I\widetilde H)\,d^{cJ}
  - (Y_e)_{IJ}\,(L_L^I\widetilde H)\,e^{cJ}
  + \text{Hermitian conjugate},
\]
where \(I,J=1,2,3\) are generation labels and all gauge indices are contracted
by the invariant tensors of \(SU(3)_c\) and \(SU(2)_L\).  In components,
\((QH)u^c\) abbreviates the contraction of the two Weyl spinor indices by
\(\epsilon_{\alpha\beta}\), the two weak doublet indices by
\(\epsilon_{ij}\), and the color indices by the natural
\(\mathbf3\otimes\overline{\mathbf3}\to\mathbf1\) pairing.  The displayed
\(Y_u,Y_d,Y_e\) are complex \(3\times3\) matrices of local coupling
coordinates.
\end{definition}

\begin{definition}[Local Standard Model charge and bundle data]
\label{def:local-standard-model-charge-bundle-data}
The local Standard Model charge algebra is
\[
  \mathfrak g_{\rm SM}
  =
  \mathfrak{su}(3)_c\oplus\mathfrak{su}(2)_L\oplus\mathfrak u(1)_Y .
\]
The corresponding covering compact group for local representation and bundle
bookkeeping is
\[
  \widetilde G_{\rm SM}=SU(3)_c\times SU(2)_L\times U(1)_y ,
\]
where \(y:=6Y\) is the integer hypercharge coordinate for the compact
\(U(1)\) factor.  The Lie-algebra generator appearing in the electroweak
covariant derivative is still \(Y=y/6\).
A representation-theoretic global form is a quotient
\[
  G_{\rm SM}=\widetilde G_{\rm SM}/\Gamma ,
  \qquad
  \Gamma\subset Z(\widetilde G_{\rm SM}),
\]
for which the local fields in
Definitions~\ref{def:sm-one-generation-left-weyl} and
\ref{def:sm-higgs-yukawa-sector} descend to representations of
\(G_{\rm SM}\).  On a spin four-manifold \(M\), let \(S_L\) denote the
left-handed spinor bundle and let \(E_R:=P\times_{G_{\rm SM}}R\) be the vector
bundle associated to a representation \(R\).  The local classical field data
consist of \(P\), a connection on \(P\), left-Weyl fields in
\(\Gamma(S_L\otimes E_R)\) for the three generations of the representations
in Definition~\ref{def:sm-one-generation-left-weyl}, a Higgs field in
\(\Gamma(E_H)\) for \(H\in(\mathbf1,\mathbf2)_{1/2}\), and the local
Yang--Mills, scalar, and Yukawa action.  The perturbative electroweak quantum
theory further requires gauge fixing or BV data, a subtraction scheme, and the
renormalized construction of Green functions and operator insertions as formal
power series to all orders in the electroweak couplings.  A cutoff regulator
may be introduced in a proof or computation, but then it is an auxiliary
regularization device to be removed or matched away; it is not the definition
of the electroweak sector and it is not a parameter in the Standard Model
prediction.
\end{definition}

The product \(\widetilde G_{\rm SM}\) in
Definition~\ref{def:local-standard-model-charge-bundle-data} is not, by itself, the
gauge group of a completed quantum Standard Model.  In the hybrid definition
of Definition~\ref{def:hybrid-standard-model-definition}, the \(SU(3)_c\) factor
belongs to the nonperturbative QCD Wightman component, while
\(SU(2)_L\times U(1)_Y\) belongs to the all-orders renormalized perturbative
electroweak component; their coupling is mediated by the operator coupling
\(\mathcal C_{\rm QCD\text{-}EW}\).  The product notation records the common
local charge assignments, anomaly sums, and possible bundle/global-form
bookkeeping that these components must share.  The hybrid definition treats the
color sector, the electroweak perturbative sector, and their coupling map as
distinct pieces of structure rather than as one declared nonperturbative path
integral with all gauge factors on the same footing.

The quotient \(\Gamma\) is invisible in the Lie algebra and in local
perturbative vertices, but it changes the allowed bundles, line operators,
and higher-form backgrounds in any formulation where those objects are
defined.  Those global questions belong to the generalized-symmetry analysis
in Chapter~\ref{ch:global-forms-higher-form-symmetry}.  This chapter uses
only the local representation-theoretic data unless a bundle or line operator
is explicitly mentioned.

\begin{remark}[Electroweak perturbation theory and optional cutoffs]
\label{rem:sm-cutoff-path-integral-not-baseline}
The baseline electroweak component of the hybrid Standard Model is
\(\mathfrak D_{\rm EW}^{\rm pert}\): an all-orders renormalized perturbative
chiral gauge theory, formulated after gauge fixing or in BV language, with a
specified subtraction and composite-operator scheme.  It is not a
finite-cutoff theory.  In a proof or computation one may introduce a
cutoff-like parameter for electroweak fields, external sources, or a weak
effective theory; such a parameter is then auxiliary regularization data or
extra Wilsonian/EFT data for a different problem, not a hidden ultraviolet
cutoff in the Standard Model prediction framework.  The color sector is not
thereby made into a cutoff sector of the same path integral; it is supplied
through the QCD component \(\mathfrak D_{\rm QCD}^{\rm W}\), with lattice or
other nonperturbative regulators used only to approximate the required
gauge-invariant QCD matrix elements.

One could instead pose a separate formal construction problem in which the QCD
sector is kept as a continuum Wightman theory while an ultraviolet regulator
is placed only on electroweak perturbative fields and sources.  Such a
construction would have to define mixed generating functionals whose
coefficients are QCD Wightman distributions of the gauge-invariant operators
in Definition~\ref{def:sm-qcd-ew-coupling-data}; it would also have to
specify the extension of those distributions at coincident points, the
contact-term scheme, and the electroweak Ward or BV identities at finite
electroweak cutoff.  This is a coherent possible formalization of the hybrid
viewpoint, but it is additional mathematical data beyond
\(\mathfrak D_{\rm EW}^{\rm pert}\), it is not the usual textbook path
integral, and it is not ordinarily how Standard Model phenomenological
computations are presented.

This is different from postulating a single ultraviolet-regulated path
integral for the full local product data
\(SU(3)_c\times SU(2)_L\times U(1)_Y\).  Such a full cutoff formulation would
have to regulate the \(SU(3)_c\) gauge fields and the chiral electroweak
fields in one coupled object, choose the phase of the chiral determinant,
provide reflection or reconstruction data, and formulate a continuum-limit
problem for the coupled theory.  That stronger object is the kind of data
discussed later in
Definition~\ref{def:sm-finite-chiral-gauge-regulator-data}; it is not the
baseline Standard Model prediction framework used in this chapter.
\end{remark}

\begin{proposition}[Kernel of the minimal Standard Model representations]
\label{prop:sm-zsix-global-form}
Let \(y:=6Y\) be the integer hypercharge coordinate, so the minimal fields
have
\[
  y(Q_L,u^c,d^c,L_L,e^c,H)=(1,-4,2,-3,6,3).
\]
Let \(\omega_3=\exp(2\pi\ii/3)\).  The element
\[
  z=\bigl(\omega_3,-1,\exp(\ii\pi/3)\bigr)
  \in SU(3)_c\times SU(2)_L\times U(1)_y
  \label{eq:sm-zsix-generator}
\]
acts trivially on every field in
Definitions~\ref{def:sm-one-generation-left-weyl} and
\ref{def:sm-higgs-yukawa-sector}.  It generates a cyclic subgroup
\(\mathbb Z_6\) of the center, and the minimal local field content descends
through any quotient by a subgroup of this \(\mathbb Z_6\):
\[
  \Gamma=1,\quad \mathbb Z_2,\quad \mathbb Z_3,\quad \mathbb Z_6 .
\]
\end{proposition}

\begin{proof}
For a representation with color triality \(t\in\mathbb Z/3\mathbb Z\), weak
doublet parity \(s\in\mathbb Z/2\mathbb Z\), and integer hypercharge \(y\),
the element \(z\) acts by the phase
\[
  \exp\!\left[
    2\pi\ii\left(\frac{t}{3}+\frac{s}{2}+\frac{y}{6}\right)
  \right].
\]
For the six minimal representations the triples \((t,s,y)\) are
\[
\begin{array}{c|c}
  Q_L & (1,1,1)\\
  u^c & (-1,0,-4)\\
  d^c & (-1,0,2)\\
  L_L & (0,1,-3)\\
  e^c & (0,0,6)\\
  H & (0,1,3).
\end{array}
\]
In each row \(t/3+s/2+y/6\) is an integer.  Thus \(z\) acts trivially.
The element has order six, because its three center components have orders
three, two, and six.  Since every subgroup of a trivially acting central
subgroup also acts trivially, the four displayed quotients are allowed by the
minimal representation content.  Distinguishing these global forms requires
line-operator and bundle data, not only the local Lie algebra.
\end{proof}

\begin{proposition}[Gauge invariance of the minimal Yukawa sector]
\label{prop:sm-yukawa-gauge-invariance}
The three terms in Definition~\ref{def:sm-higgs-yukawa-sector} are precisely
the gauge-invariant local monomials that are bilinear in the listed
left-handed Weyl fields, linear in \(H\) or \(\widetilde H\), and of
engineering dimension at most four, up to flavor tensors and Hermitian
conjugation.
\end{proposition}

\begin{proof}
Lorentz invariance forces the two left-handed Weyl spinor indices in each
fermion bilinear to be contracted by the antisymmetric invariant tensor of
\(SL(2,\mathbb C)\).  Color invariance requires \(Q_L\) to pair with either
\(u^c\) or \(d^c\), because the quark singlet contraction is
\(\mathbf3\otimes\overline{\mathbf3}\to\mathbf1\); the lepton bilinear is
already color neutral.  Weak \(SU(2)\)-invariance then forces the remaining
weak doublet index to be contracted with either \(H\) or \(\widetilde H\) by
the invariant tensor of the doublet representation.

The hypercharge sums determine which Higgs doublet occurs:
\[
  \frac16+\frac12-\frac23=0,\qquad
  \frac16-\frac12+\frac13=0,\qquad
  -\frac12-\frac12+1=0 .
\]
Thus \(Q_LHu^c\), \(Q_L\widetilde H d^c\), and
\(L_L\widetilde H e^c\) are gauge invariant.  The opposite choices of
\(H\) and \(\widetilde H\) have nonzero hypercharge, hence do not define
local gauge-invariant monomials in the left-handed convention.  Power
counting excludes additional renormalizable matter monomials with more fields
or derivatives in this minimal field content.  Flavor indices are spectators
for the local charge algebra, so the allowed coefficients are the three
arbitrary complex flavor matrices shown above.
\end{proof}

Flavor rotations of the fermion fields remove unphysical parameters; the
remaining charged-current quark mixing is encoded in the CKM matrix.  If
neutrino masses are included through dimension-five or singlet-neutrino
operators, a separate PMNS mixing matrix appears in the lepton sector.

\section{Flavor Coordinates and Charged Currents}

The Yukawa matrices are not themselves the physical flavor parameters.  A
physical statement must specify which field redefinitions have already been
used and which current is being observed.

\begin{definition}[Quark mass coordinates and CKM matrix]
\label{def:sm-ckm-matrix}
After evaluating the Higgs field at \(H_0\), define the complex quark mass
matrices
\[
  M_u=\frac{v}{\sqrt2}Y_u,\qquad
  M_d=\frac{v}{\sqrt2}Y_d .
\]
Choose singular-value decompositions
\[
  (U_L^u)^\dagger M_uU_R^u=D_u,\qquad
  (U_L^d)^\dagger M_dU_R^d=D_d ,
  \label{eq:sm-quark-svd}
\]
where \(U_L^u,U_R^u,U_L^d,U_R^d\in U(3)\) and \(D_u,D_d\) are diagonal with
nonnegative entries.  The Cabibbo--Kobayashi--Maskawa matrix is
\[
  V_{\rm CKM}:=(U_L^u)^\dagger U_L^d .
  \label{eq:sm-ckm-definition}
\]
It is the coefficient matrix in the charged weak current written in quark
mass coordinates:
\[
  J_\mu^+
  =
  \overline u_{L,a}\gamma_\mu
  (V_{\rm CKM})_{ab}d_{L,b},
  \qquad
  J_\mu^-=(J_\mu^+)^\dagger .
\]
Here \(u_L,d_L\) denote mass-basis left-handed quark fields.  The neutral
gauge currents remain diagonal in flavor in these coordinates.
\end{definition}

\begin{proposition}[Flavor parameter counting]
\label{prop:sm-flavor-parameter-counting}
For \(n\) quark generations with nondegenerate masses, the pair of complex
Yukawa matrices \((Y_u,Y_d)\) contains
\[
  2n \quad\text{quark masses},\qquad
  \frac{n(n-1)}{2}\quad\text{mixing angles},\qquad
  \frac{(n-1)(n-2)}{2}\quad\text{CP-odd phases}.
\]
In particular, two generations allow one mixing angle and no CKM CP phase,
while three generations allow three angles and one CKM CP phase.
\end{proposition}

\begin{proof}
The two complex \(n\times n\) Yukawa matrices contain \(4n^2\) real
parameters.  In the absence of Yukawa couplings the quark kinetic and gauge
terms have flavor symmetry
\[
  U(n)_Q\times U(n)_u\times U(n)_d .
\]
The Yukawa matrices break this group to the diagonal baryon-number
\(U(1)_B\), which does not change \(Y_u\) or \(Y_d\).  Therefore
\(3n^2-1\) real parameters are removable by flavor-coordinate choices.
The physical quark-sector parameter count is
\[
  4n^2-(3n^2-1)=n^2+1 .
\]
Among these are \(2n\) nonnegative singular values, namely the quark masses.
Thus the mixing matrix carries
\[
  n^2+1-2n=(n-1)^2
\]
parameters.  A unitary \(n\times n\) matrix has \(n^2\) real parameters.
After the masses have been diagonalized, the \(2n\) quark mass eigenfields
may be rephased, but the common baryon-number rephasing leaves
\(V_{\rm CKM}\) unchanged.  Hence \(2n-1\) phases are unphysical in \(V_{\rm
CKM}\).  The remaining \((n-1)^2\) parameters split into
\(n(n-1)/2\) real rotation angles and
\[
  (n-1)^2-\frac{n(n-1)}{2}=\frac{(n-1)(n-2)}{2}
\]
phases.
\end{proof}

\begin{proposition}[Tree-level GIM diagonality]
\label{prop:sm-gim-tree-level-diagonality}
In the quark mass basis of Definition~\ref{def:sm-ckm-matrix}, the neutral
electromagnetic, gluon, and \(Z\)-boson currents are flavor diagonal at tree
level.  Flavor change in the renormalizable gauge currents is confined to the
charged current and is controlled by \(V_{\rm CKM}\).
\end{proposition}

\begin{proof}
The gluon and photon couple to generators proportional to the identity in
generation space.  A unitary change of mass coordinates therefore gives
\[
  \overline q_L U_L^\dagger\gamma_\mu T U_L q_L
  =
  \overline q_L\gamma_\mu T q_L
\]
whenever \(T\) acts only on color or electric charge and not on generation.
The \(Z\)-current is a linear combination of \(T_L^3\) and \(Q_{\rm em}\).
Within the up sector and within the down sector these charges are again
generation-independent, so the same unitarity cancellation applies.  The
charged current is different because it pairs the up component of \(Q_L\)
with the down component before the two independent left singular-vector
matrices have been chosen.  Its flavor matrix is therefore
\((U_L^u)^\dagger U_L^d=V_{\rm CKM}\).
\end{proof}

\begin{definition}[Jarlskog invariant]
\label{def:sm-jarlskog-invariant}
For three generations, a rephasing-invariant CKM CP-odd coordinate is
\[
  J_{\rm CKM}
  =
  \operatorname{Im}
  \left(
    V_{us}V_{cb}V_{ub}^*V_{cs}^*
  \right).
\]
Equivalently, any expression
\(\operatorname{Im}(V_{ij}V_{kl}V_{il}^*V_{kj}^*)\) with distinct
\(i,k\) and distinct \(j,l\) is a rephasing-invariant quartet.  For a
three-by-three unitary matrix these quartets are related by unitarity and
carry the same magnitude, with signs fixed by row and column orientation.
\end{definition}

\begin{proposition}[Rephasing invariance of \(J_{\rm CKM}\)]
\label{prop:sm-jarlskog-rephasing}
Under mass-basis quark rephasings
\[
  V_{ij}\longmapsto \ee^{\ii\alpha_i}V_{ij}\ee^{-\ii\beta_j},
\]
the quantity \(J_{\rm CKM}\) is invariant.  For \(n=2\) the corresponding
quartet invariant vanishes after rephasing, so a two-generation charged
current has no intrinsic CKM CP phase.
\end{proposition}

\begin{proof}
The phase of the quartet
\[
  V_{us}V_{cb}V_{ub}^*V_{cs}^*
\]
is multiplied by
\[
  \ee^{\ii\alpha_u}\ee^{-\ii\beta_s}
  \ee^{\ii\alpha_c}\ee^{-\ii\beta_b}
  \ee^{-\ii\alpha_u}\ee^{\ii\beta_b}
  \ee^{-\ii\alpha_c}\ee^{\ii\beta_s}=1 .
\]
Thus its imaginary part is a coordinate-independent CP-odd function of the
charged current.  The equality of magnitudes of the different three-generation
quartets follows directly from row and column orthogonality.  For example,
if \(i\neq k\) and \(j,l,n\) are the three column labels, row orthogonality
gives
\[
  V_{ij}V_{kj}^*+V_{il}V_{kl}^*+V_{in}V_{kn}^*=0 .
\]
Multiplying by \(V_{kl}V_{il}^*\) and taking imaginary parts gives
\[
  \operatorname{Im}(V_{ij}V_{kl}V_{il}^*V_{kj}^*)
  =
  -\operatorname{Im}(V_{in}V_{kl}V_{il}^*V_{kn}^*) .
\]
Column orthogonality gives the analogous relation for fixed column pair and
varying rows.  Combining the row and column relations moves any quartet with
distinct row labels and distinct column labels to any other such quartet;
interchanging the ordered row pair or the ordered column pair changes the
sign.  For two generations,
Proposition~\ref{prop:sm-flavor-parameter-counting} shows that all phases in
a \(2\times2\) unitary charged-current matrix are removable by quark-field
rephasings.  A representative can therefore be chosen real, and every
quartet imaginary part is zero in that coordinate.
\end{proof}

\section{Electroweak Breaking and Mass Coordinates}

The word ``breaking'' in this section refers to a local field coordinate
around a Higgs background and to the stabilizer of that background.  Gauge
redundancy is not a global symmetry acting faithfully on the physical Hilbert
space, and it is not literally broken.  The physical content of the
Brout--Englert--Higgs phase is the spectrum, currents, and gauge-invariant
operator algebra described in the renormalized perturbative construction
after gauge fixing or BV treatment, or in a nonperturbative construction if
one has been supplied.  The formulas below are the tree-level mass and
coupling coordinates of that local electroweak chart.

\begin{definition}[Electroweak scalar chart]
\label{def:sm-electroweak-scalar-chart}
The minimal renormalizable electroweak scalar potential is
\[
  V(H)=\lambda\left(H^\dagger H-\frac{v^2}{2}\right)^2 ,
  \qquad \lambda>0 .
\]
In the local field-space chart used below,
\[
  H(x)=\frac1{\sqrt2}
  \begin{pmatrix}
    \chi_1(x)+\ii\chi_2(x)\\
    v+h(x)+\ii\chi_3(x)
  \end{pmatrix}.
\]
The reference vacuum vector is
\[
  H_0=\frac1{\sqrt2}\begin{pmatrix}0\\ v\end{pmatrix}
\]
in the same chart.  Here \(v>0\) and \(\lambda>0\) are renormalized
coordinates in the perturbative electroweak construction.  The fields
\(\chi_1,\chi_2,\chi_3\) are local coordinates tangent to the gauge orbit
through \(H_0\), while \(h\) is the radial scalar coordinate.
\end{definition}

\begin{proposition}[Electromagnetic stabilizer and weak-boson masses]
\label{prop:sm-electroweak-stabilizer-masses}
In the convention of
Convention~\ref{conv:sm-electroweak-trace-normalization}, the stabilizer of
\(H_0\) inside the local group \(SU(2)_L\times U(1)_Y\) is generated by
\[
  Q_{\rm em}=T^3_L+Y .
\]
The quadratic gauge-boson mass coordinates are
\[
  W^\pm_\mu=\frac1{\sqrt2}(\mathcal W^1_\mu\mp\ii\mathcal W^2_\mu),
  \qquad
  Z_\mu=\frac{g_2\mathcal W^3_\mu-g_1\mathcal B_\mu}
              {\sqrt{g_1^2+g_2^2}},
  \qquad
  A_\mu=\frac{g_1\mathcal W^3_\mu+g_2\mathcal B_\mu}
              {\sqrt{g_1^2+g_2^2}},
\]
with
\[
  m_W=\frac{g_2v}{2},
  \qquad
  m_Z=\frac{v}{2}\sqrt{g_1^2+g_2^2},
  \qquad
  m_A=0 .
\]
The coupling of the massless field is
\[
  e=g_2\sin\theta_W=g_1\cos\theta_W,
  \qquad
  \tan\theta_W=\frac{g_1}{g_2}.
\]
\end{proposition}

\begin{proof}
The weak generator \(T^3_L=\operatorname{diag}(1/2,-1/2)\) acts on
\(H_0\) by \(-H_0/2\), while the hypercharge generator acts by
\(H_0/2\).  Hence \((T^3_L+Y)H_0=0\).  The generators \(T^1_L\) and
\(T^2_L\) move \(H_0\) into the upper component, and a general combination
\(aT^3_L+bY\) annihilates \(H_0\) only when \(a=b\).  Therefore the
one-dimensional stabilizer Lie algebra is spanned by \(Q_{\rm em}\).

The covariant derivative in the explicit-coupling convention is
\[
  D_\mu H
  =
  \left(
    \partial_\mu-\ii g_2\mathcal W^a_\mu\tau^a-\ii g_1\mathcal B_\mu Y
  \right)H .
\]
Using
\[
  \tau^1H_0=\frac{v}{2\sqrt2}\begin{pmatrix}1\\0\end{pmatrix},
  \qquad
  \tau^2H_0=\frac{v}{2\sqrt2}\begin{pmatrix}-\ii\\0\end{pmatrix},
  \qquad
  \tau^3H_0=-\frac12H_0,
  \qquad
  YH_0=\frac12H_0 ,
\]
the quadratic term from \((D_\mu H)^\dagger(D^\mu H)\) at \(H=H_0\) is
\[
  (D_\mu H_0)^\dagger(D^\mu H_0)
  =
  \frac{v^2}{8}
  \left[
    g_2^2(\mathcal W^1_\mu\mathcal W^{1\mu}
          +\mathcal W^2_\mu\mathcal W^{2\mu})
    +(-g_2\mathcal W^3_\mu+g_1\mathcal B_\mu)^2
  \right].
\]
The charged part is
\((g_2^2v^2/4)W^+_\mu W^{-\mu}\), so
\(m_W^2=g_2^2v^2/4\).  The neutral quadratic form has rank one: the massive
unit vector is \(Z_\mu\), the null vector is \(A_\mu\), and comparison with
the real-vector mass term \(\frac12m_Z^2Z_\mu Z^\mu\) gives
\(m_Z^2=(g_1^2+g_2^2)v^2/4\).

Finally invert the neutral rotation:
\[
  \mathcal W^3_\mu
  =
  \frac{g_2Z_\mu+g_1A_\mu}{\sqrt{g_1^2+g_2^2}},
  \qquad
  \mathcal B_\mu
  =
  \frac{-g_1Z_\mu+g_2A_\mu}{\sqrt{g_1^2+g_2^2}} .
\]
The coefficient of \(A_\mu\) in
\(g_2\mathcal W^3_\mu T^3_L+g_1\mathcal B_\mu Y\) is
\[
  \frac{g_1g_2}{\sqrt{g_1^2+g_2^2}}\,(T^3_L+Y).
\]
This proves the displayed electromagnetic coupling and identifies the
massless gauge field with the connection for the unbroken stabilizer.
\end{proof}

This is the nonabelian mass-matrix construction of
Chapter~\ref{ch:vol2-classical-yang-mills-matter} applied to the
\((\mathbf2)_{1/2}\) scalar representation.  The three coordinates
\(\chi_1,\chi_2,\chi_3\) are gauge-orbit coordinates in a local chart; in an
\(R_\xi\) gauge they remain as unphysical fields paired by BRST, while in a
unitary-gauge chart they are removed locally.

\paragraph{Higgs radial mass coordinate.}
In the scalar chart of Definition~\ref{def:sm-electroweak-scalar-chart}, the
potential restricted to the radial coordinate is
\[
  V(h)=\lambda\left(vh+\frac{h^2}{2}\right)^2
  =
  \lambda v^2h^2+\lambda vh^3+\frac{\lambda}{4}h^4 .
\]
Consequently the radial scalar mass coordinate is
\[
  m_h^2=2\lambda v^2 .
\]
The cubic and quartic self-interaction coefficients in the potential are
\(\lambda v\) and \(\lambda/4\), respectively, in this field coordinate.
In the unitary local chart \(H=(0,(v+h)/\sqrt2)^T\),
\[
  H^\dagger H-\frac{v^2}{2}=vh+\frac{h^2}{2}.
\]
Substitution into \(V(H)\) gives the displayed polynomial.  The Lorentzian
Lagrangian contains \(-V\), while the standard real-scalar mass term is
\(-\frac12m_h^2h^2\).  Comparing the quadratic coefficients gives
\(m_h^2=2\lambda v^2\).

\begin{proposition}[Custodial \(SU(2)\) and the tree-level \(\rho\) coordinate]
\label{prop:sm-custodial-rho}
Set
\[
  \Sigma:=(\widetilde H,H).
\]
The scalar kinetic term with \(g_1=0\) and the scalar potential are invariant
under
\[
  \Sigma\longmapsto U_L\Sigma U_R^\dagger,
  \qquad U_L,U_R\in SU(2),
\]
and the vacuum \(\Sigma_0=(v/\sqrt2)\mathbf1_2\) preserves the diagonal
subgroup \(SU(2)_V\).  With hypercharge gauged as the \(T^3_R\) subgroup, the
tree-level mass ratio
\[
  \rho_{\rm tree}
  :=
  \frac{m_W^2}{m_Z^2\cos^2\theta_W}
\]
is equal to \(1\).
\end{proposition}

\begin{proof}
The identity
\[
  \frac{1}{2}\operatorname{tr}(\Sigma^\dagger\Sigma)=H^\dagger H
\]
shows that the scalar potential depends only on
\(\operatorname{tr}(\Sigma^\dagger\Sigma)\), hence is invariant under
left and right multiplication by unitary matrices.  With \(g_1=0\), the
covariant derivative acts only by left \(SU(2)_L\) multiplication, so the
right \(SU(2)_R\) remains a global symmetry of the scalar kinetic term.  The
vacuum equation \(U_L\Sigma_0U_R^\dagger=\Sigma_0\) is equivalent to
\(U_L=U_R\), giving the diagonal stabilizer.

The hypercharge gauge field gauges only the \(T^3_R\) direction and therefore
breaks the global \(SU(2)_R\) explicitly, but the mass formulae of
Proposition~\ref{prop:sm-electroweak-stabilizer-masses} still give
\[
  \cos\theta_W=\frac{g_2}{\sqrt{g_1^2+g_2^2}},\qquad
  m_W^2=\frac{g_2^2v^2}{4},\qquad
  m_Z^2=\frac{(g_1^2+g_2^2)v^2}{4}.
\]
Substitution gives \(\rho_{\rm tree}=1\).  Radiative corrections and
higher-dimensional operators are separate coordinates of the electroweak
EFT; they are not part of this tree-level identity.
\end{proof}

\section{Gauge-Anomaly Cancellation}

\begin{definition}[Gauge-anomaly coefficients for the local field content]
\label{def:sm-local-anomaly-coefficients}
For a collection of left-handed Weyl fermions in representations \(R_i\) of
\(\mathfrak g_{\rm SM}\), the perturbative gauge-anomaly tensor is the
symmetric cubic functional
\[
  \mathcal A(X,Y,Z)
  =
  \sum_i \operatorname{tr}_{R_i}\bigl(X\{Y,Z\}\bigr),
  \qquad
  X,Y,Z\in\mathfrak g_{\rm SM},
\]
with spectator multiplicities included in the trace.  For a mixed
\(G^2U(1)_Y\) component, this reduces to the sum of hypercharge times the
quadratic index of the \(G\)-representation.  For the mixed
gravitational-\(U(1)_Y\) anomaly, the coefficient is the sum of hypercharges
over all left-handed Weyl components, again including spectator
multiplicities.
\end{definition}

\begin{proposition}[SM anomaly cancellation]
\label{prop:sm-generation-anomaly-cancellation}
For the one-generation representation of
Definition~\ref{def:sm-one-generation-left-weyl}, all perturbative local
gauge anomalies and the mixed gravitational-hypercharge anomaly vanish.  The
finite \(SU(2)_L\) anomaly also vanishes generation by generation.
\end{proposition}

\begin{proof}
The calculation uses only the representation table and the fact that all
fermions have been written as left-handed Weyl fields.  For the pure
\(SU(3)_c^3\) anomaly, the two weak components of \(Q_L\) contribute two
fundamental color representations, while \(u^c\) and \(d^c\) contribute two
antifundamentals.  Since the cubic index changes sign under complex
conjugation,
\[
  2A(\mathbf3)-A(\mathbf3)-A(\mathbf3)=0 .
\]
The pure \(SU(2)_L^3\) perturbative anomaly vanishes because the symmetric
cubic invariant \(d^{abc}\) of \(\mathfrak{su}(2)\) is zero.  Mixed
\(SU(3)_cU(1)_Y^2\) and \(SU(2)_LU(1)_Y^2\) components vanish because they
contain \(\operatorname{tr}_{R}(T^a)\) for a semisimple generator.

For \(SU(3)_c^2U(1)_Y\), one generation gives
\[
  \mathcal A[SU(3)^2U(1)_Y]
  =
  2\left(\frac16\right)T(\mathbf3)
  +\left(-\frac23\right)T(\overline{\mathbf3})
  +\left(\frac13\right)T(\overline{\mathbf3})
  =0 .
\]
Here \(T(\overline{\mathbf3})=T(\mathbf3)\), and the factor \(2\) is the weak
doublet multiplicity.  For \(SU(2)_L^2U(1)_Y\),
\[
  \mathcal A[SU(2)^2U(1)_Y]
  =
  3\left(\frac16\right)T(\mathbf2)
  +\left(-\frac12\right)T(\mathbf2)
  =0 .
\]
Here the factor \(3\) is the color multiplicity of \(Q_L\).  For the cubic
hypercharge anomaly, include the full dimensions of the nonabelian
representations:
\[
\begin{aligned}
  \mathcal A[U(1)_Y^3]
  &=
  6\left(\frac16\right)^3
  +3\left(-\frac23\right)^3
  +3\left(\frac13\right)^3
  +2\left(-\frac12\right)^3
  +1^3\\
  &=0 .
\end{aligned}
\]
The mixed gravitational-hypercharge anomaly coefficient is
\[
  6\left(\frac16\right)
  +3\left(-\frac23\right)
  +3\left(\frac13\right)
  +2\left(-\frac12\right)
  +1
  =0 .
\]
Finally, the finite \(SU(2)\) anomaly of
Chapter~\ref{ch:volume-ii-21} is absent because each
generation contains four left-handed \(SU(2)_L\) doublets: three colored
quark doublets and one lepton doublet.  The number of doublets is therefore
even.  These equalities are finite representation-theoretic constraints on
the regulated fermion determinant line; they are not consequences of
perturbative renormalizability alone.
\end{proof}

\section{Baryon and Lepton Number as Global Data}

\begin{definition}[Classical \(B\), \(L\), and \(B-L\) charges]
\label{def:sm-b-l-charges}
In the left-handed Weyl convention, assign baryon and lepton charges
\[
\begin{array}{c|cc}
  \text{field} & B & L \\ \hline
  Q_L & 1/3 & 0\\
  u^c & -1/3 & 0\\
  d^c & -1/3 & 0\\
  L_L & 0 & 1\\
  e^c & 0 & -1\\
  \nu^c\text{, if included} & 0 & -1 .
\end{array}
\]
The signs for conjugate fields are the charges of the left-handed conjugate
fields themselves.  These assignments are statements about global-current
coordinates in the local Lagrangian; whether a global symmetry survives in
the quantum theory is determined by its anomaly polynomial and by possible
operator insertions.
\end{definition}

\begin{proposition}[\(B+L\) anomaly and \(B-L\) mixed-gauge cancellation]
\label{prop:sm-baryon-lepton-anomaly}
For one generation of the minimal representation without \(\nu^c\), the
mixed \(SU(2)_L^2U(1)_B\) and \(SU(2)_L^2U(1)_L\) anomaly coefficients are
both nonzero:
\[
  \mathcal A[SU(2)_L^2U(1)_B]=T(\mathbf2),\qquad
  \mathcal A[SU(2)_L^2U(1)_L]=T(\mathbf2).
\]
Their difference vanishes, so \(B-L\) has no mixed
\(SU(2)_L^2\) anomaly.  The mixed \(SU(3)_c^2U(1)_{B-L}\) anomaly also
vanishes.  The gravitational and cubic anomalies of \(B-L\) vanish
generation by generation only after a gauge-singlet \(\nu^c\) with
\(B-L=1\) is included.
\end{proposition}

\begin{proof}
The \(SU(2)^2U(1)\) coefficient is the sum of the global charge times the
quadratic index of each weak doublet, with color multiplicity included.  For
baryon number only \(Q_L\) contributes, giving
\[
  3\left(\frac{1}{3}\right)T(\mathbf2)=T(\mathbf2).
\]
For lepton number only \(L_L\) contributes, giving \(T(\mathbf2)\).  Thus
\(B+L\) is anomalous and \(B-L\) cancels in the \(SU(2)^2\) channel.  For
\(SU(3)^2(B-L)\), the common \(T(\mathbf3)\) factors out and the coefficient
is
\[
  2\left(\frac{1}{3}\right)-\frac{1}{3}-\frac{1}{3}=0,
\]
where the factor \(2\) is the weak multiplicity of \(Q_L\).

For the mixed gravitational anomaly of \(B-L\), including multiplicities but
not \(\nu^c\), one obtains
\[
  6\left(\frac{1}{3}\right)
  +3\left(-\frac{1}{3}\right)
  +3\left(-\frac{1}{3}\right)
  +2(-1)+1=-1 .
\]
The cubic \(U(1)_{B-L}^3\) coefficient gives the same value:
\[
  6\left(\frac{1}{3}\right)^3
  +3\left(-\frac{1}{3}\right)^3
  +3\left(-\frac{1}{3}\right)^3
  +2(-1)^3+1^3=-1 .
\]
Adding a singlet \(\nu^c\) with \(B-L=1\) contributes \(+1\) to each of
these two anomaly coefficients and cancels them.  This is a finite anomaly
calculation, not a statement that a singlet neutrino has been observed or that
one has chosen a particular neutrino-mass mechanism.
\end{proof}

\begin{remark}[Sphalerons as transitions between electroweak sectors]
The \(SU(2)_L\) anomaly means that an electroweak gauge-field history with
topological charge \(Q_{\rm top}\) changes the global charges by
\(\Delta B=\Delta L=N_gQ_{\rm top}\), where \(N_g\) is the number of
generations, while \(\Delta(B-L)=0\).  This statement assumes a field history
whose boundary conditions make the \(SU(2)\) second Chern number an integer;
on more general backgrounds the corresponding statement must be formulated in
terms of the anomaly inflow class.  The sphaleron is the unstable static
configuration at the top of the minimal barrier between neighboring
electroweak sectors, as described in
Chapter~\ref{ch:vol2-classical-yang-mills-matter}.  Rates in a thermal state
require real-time finite-temperature dynamics and are not fixed by the
topological selection rule alone.
\end{remark}

\section{Higher-Dimensional Operator Coordinates}

\begin{definition}[SMEFT coupling chart]
\label{def:sm-smeft-coupling-chart}
Fix a regulator, a local operator basis, and a subtraction scheme.  The
corresponding Standard-Model effective coupling chart below a scale
\(\Lambda\) is the local functional
\[
  S_{\rm EFT}
  =
  S_{\rm SM}^{\rm ren}
  +
  \sum_{d>4}\Lambda^{4-d}
  \sum_{A\in\mathcal B_d} C_A^{(d)}(\mu)
  \int_{M_4}\mathcal O_A^{(d)}(\mu),
  \label{eq:sm-smeft-coupling-chart}
\]
where \(\mathcal B_d\) is a specified basis of gauge-invariant local
operators of engineering dimension \(d\), after quotienting by total
derivatives and by the chosen field-redefinition equations of motion.  The
symbol \(\Lambda\) is a reference scale used to write dimensionless Wilson
coordinates in an effective theory beyond the renormalizable Standard Model.
It is not a hidden cutoff parameter of Standard Model predictions.  If the
same expansion is interpreted as a Wilsonian theory with a physical cutoff,
that interpretation is additional regulator and matching data.  The
coordinates \(C_A^{(d)}(\mu)\) are not observables by themselves; an
observable uses them together with the operator scheme, matching map, field
domain, and matrix elements.
\end{definition}

\begin{definition}[Weinberg operator in the left-handed convention]
\label{def:sm-weinberg-operator}
With only the minimal fields of
Definitions~\ref{def:sm-one-generation-left-weyl} and
\ref{def:sm-higgs-yukawa-sector}, the derivative-free dimension-five
lepton-number-violating operator is
\[
  \mathcal O_5^{IJ}
  =
  \epsilon^{\alpha\beta}
  \bigl(\epsilon_{ij}L^I_{\alpha i}H_j\bigr)
  \bigl(\epsilon_{k\ell}L^J_{\beta k}H_\ell\bigr),
  \label{eq:sm-weinberg-operator}
\]
with symmetric flavor coefficient \(C_5^{IJ}=C_5^{JI}\).  In the convention
\[
  \Delta\mathcal L_5
  =
  -\frac{1}{2\Lambda}C_5^{IJ}\mathcal O_5^{IJ}
  +\text{Hermitian conjugate},
  \label{eq:sm-weinberg-lagrangian-convention}
\]
the neutral-lepton Majorana mass matrix in the electroweak scalar chart is
\[
  (M_\nu)^{IJ}=\frac{v^2}{2\Lambda}C_5^{IJ}.
  \label{eq:sm-weinberg-neutrino-mass}
\]
\end{definition}

To verify \eqref{eq:sm-weinberg-neutrino-mass}, set \(H=H_0\).  The weak
contraction
\(\epsilon_{ij}L^I_iH_j\) selects the neutral component of \(L^I\) and gives
\[
  \epsilon_{ij}L^I_{\alpha i}H_{0,j}
  =
  \frac{v}{\sqrt2}\,\nu^I_{\alpha}.
\]
Substitution into \eqref{eq:sm-weinberg-lagrangian-convention} gives
\[
  \Delta\mathcal L_5
  =
  -\frac{1}{2}
  \left(\frac{v^2}{2\Lambda}C_5^{IJ}\right)
  \epsilon^{\alpha\beta}\nu^I_\alpha\nu^J_\beta
  +\text{Hermitian conjugate},
\]
which is the displayed Majorana mass coordinate.  The symmetry of
\(C_5^{IJ}\) follows because the Lorentz contraction of two left-handed Weyl
fields is symmetric in the flavor labels after the anticommutation of the
fields is combined with the antisymmetry of
\(\epsilon^{\alpha\beta}\).

\begin{proposition}[Integrating out singlet neutrinos at tree level]
\label{prop:sm-type-i-seesaw-matching}
Add gauge-singlet left-handed fields \(\nu^{cA}\) with
\[
  \Delta\mathcal L
  =
  -Y_\nu^{IA}(L^I H)\nu^{cA}
  -\frac{1}{2}M_{AB}\nu^{cA}\nu^{cB}
  +\text{Hermitian conjugate},
\]
where \(M\) is a nonsingular complex symmetric matrix and spinor and weak
indices are contracted as in \eqref{eq:sm-weinberg-operator}.  At tree level
and to leading order in external momenta over the singular values of \(M\),
the dimension-five coefficient obeys
\[
  \frac{C_5}{\Lambda}
  =
  Y_\nu M^{-1}Y_\nu^{T}
\]
up to the conventional overall sign that can be changed by rephasing
Majorana fields.
\end{proposition}

\begin{proof}
The algebraic equation for the heavy singlet field, after suppressing
derivative corrections, is
\[
  M_{AB}\nu^{cB}+Y_\nu^{IA}(L^I H)=0 .
\]
Thus
\[
  \nu^{cA}=-(M^{-1})^{AB}Y_\nu^{IB}(L^I H).
\]
Substituting this solution back into the quadratic expression in
\(\nu^c\) gives a local operator quadratic in \((LH)\) with coefficient
\(\frac12Y_\nu M^{-1}Y_\nu^T\), with the sign depending on the convention
used to order the Grassmann variables in the Majorana bilinear.  Comparing
with Definition~\ref{def:sm-weinberg-operator} gives the displayed matching
coordinate.  The statement is a tree-level EFT matching computation, not a
claim that singlet neutrinos are part of the minimal Standard Model field
content.
\end{proof}

\begin{definition}[Dimension-six quotient space]
\label{def:sm-dimension-six-quotient}
Let \(\mathcal V_6\) be the complex vector space of local polynomial
gauge-invariant densities of engineering dimension six built from
\[
  G^A_{\mu\nu},\quad W^I_{\mu\nu},\quad B_{\mu\nu},\quad
  H,\quad D_\mu,\quad q,\ell,u,d,e,
\]
where \(q,\ell,u,d,e\) denote the standard chiral Standard-Model fields in
four-component notation.  This notation is a bookkeeping device for this
basis: \(q,\ell\) are the left-handed fields \(Q_L,L_L\), while \(u,d,e\)
are the right-handed conjugates of the left-handed Weyl fields
\((u^c,d^c,e^c)\) used in
Definition~\ref{def:sm-one-generation-left-weyl}.  Let
\(\partial\mathcal V_5\subset\mathcal V_6\) be total derivatives and let
\(\mathcal E_6\subset\mathcal V_6/\partial\mathcal V_5\) be the subspace
generated by
\[
  \frac{\delta S_{\rm SM}^{\rm ren}}{\delta\Phi}\,\mathcal F_\Phi
  +\text{Hermitian conjugate},
  \label{eq:sm-eom-operator-subspace}
\]
where \(\Phi\) runs over the SM fields and
\(\mathcal F_\Phi\) is any gauge-covariant local polynomial of dimension
\(6-\dim(\delta S_{\rm SM}^{\rm ren}/\delta\Phi)\).  A dimension-six
SMEFT basis is an ordered basis of the quotient
\[
  \mathcal Q_6
  :=
  \mathcal V_6/(\partial\mathcal V_5+\mathcal E_6),
  \label{eq:sm-dimension-six-quotient}
\]
together with a Hermiticity convention and a flavor-index convention for its
coefficients.
\end{definition}

\begin{proposition}[Why equation-of-motion operators are coordinates]
\label{prop:sm-eom-operator-coordinate-change}
Let \(S_{\rm SM}^{\rm ren}\) be the renormalized SM action in the chosen
regularization and subtraction scheme.  Assume the regulator treats the
near-identity local field change below as an allowed coordinate change, with
its Jacobian either trivial in dimensional regularization or absorbed into
local counterterms in a local cutoff scheme.  The infinitesimal local change of
field coordinates
\[
  \Phi\longmapsto
  \Phi+\Lambda^{-2}\mathcal F_\Phi(\Phi)
  \label{eq:sm-local-field-coordinate-change}
\]
changes the action, to order \(\Lambda^{-2}\), by an element of
\(\mathcal E_6\).  Therefore coefficients multiplying elements of
\(\mathcal E_6\) are not invariant coordinates of the effective
theory; they are removed by the quotient in
Definition~\ref{def:sm-dimension-six-quotient}.
\end{proposition}

\begin{proof}
Taylor expansion of the action as a functional gives
\[
  S_{\rm SM}^{\rm ren}[\Phi+\Lambda^{-2}\mathcal F_\Phi]
  =
  S_{\rm SM}^{\rm ren}[\Phi]
  +
  \Lambda^{-2}
  \sum_\Phi\int_{M_4}
  \frac{\delta S_{\rm SM}^{\rm ren}}{\delta\Phi(x)}
  \mathcal F_\Phi(x)
  +O(\Lambda^{-4}).
\]
Boundary terms produced by moving derivatives between
\(\delta S/\delta\Phi\) and \(\mathcal F_\Phi\) lie in
\(\partial\mathcal V_5\).  Thus the order-\(\Lambda^{-2}\) shift is exactly
of the form \eqref{eq:sm-eom-operator-subspace}.  The statement is a
coordinate statement inside a specified renormalized perturbative or regulated
effective path integral; it does not say that the corresponding composite
operator insertion vanishes as a distribution without specifying the
operator-renormalization scheme.  If the chosen scheme produces a local
Jacobian for the field redefinition, that Jacobian changes local coupling
coordinates in the same EFT chart; it does not make the equation-of-motion
operator an independent observable.
\end{proof}

\begin{definition}[Warsaw-type dimension-six basis]
\label{def:sm-warsaw-type-dimension-six-basis}
For a baryon-number preserving coupling chart with no singlet neutrino fields,
one convenient ordered basis of \(\mathcal Q_6\) is the Warsaw-type basis.
Write
\[
  \widetilde X_{\mu\nu}:=
  \frac12\epsilon_{\mu\nu\rho\sigma}X^{\rho\sigma},
  \qquad
  H^\dagger i\overleftrightarrow D_\mu H
  :=
  iH^\dagger D_\mu H-i(D_\mu H)^\dagger H,
\]
and
\[
  H^\dagger i\overleftrightarrow D_\mu^I H
  :=
  iH^\dagger\tau^I D_\mu H-i(D_\mu H)^\dagger\tau^I H .
\]
The bosonic part contains the following fifteen one-generation operator
types:
\[
\begin{aligned}
X^3&:\quad
  O_G,\ O_{\widetilde G},\ O_W,\ O_{\widetilde W},\\
H^6&:\quad
  O_H=(H^\dagger H)^3,\\
H^4D^2&:\quad
  O_{H\Box}=(H^\dagger H)\Box(H^\dagger H),\quad
  O_{HD}=|H^\dagger D_\mu H|^2,\\
X^2H^2&:\quad
  O_{HG},O_{H\widetilde G},O_{HW},O_{H\widetilde W},
  O_{HB},O_{H\widetilde B},O_{HWB},O_{H\widetilde WB}.
\end{aligned}
\]
Here
\[
\begin{aligned}
  O_G&=f^{ABC}G^A_{\mu}{}^{\nu}G^B_{\nu}{}^{\rho}
      G^C_{\rho}{}^{\mu},&
  O_{\widetilde G}&=f^{ABC}\widetilde G^A_{\mu}{}^{\nu}
      G^B_{\nu}{}^{\rho}G^C_{\rho}{}^{\mu},\\
  O_W&=\epsilon^{IJK}W^I_{\mu}{}^{\nu}W^J_{\nu}{}^{\rho}
      W^K_{\rho}{}^{\mu},&
  O_{\widetilde W}&=\epsilon^{IJK}\widetilde W^I_{\mu}{}^{\nu}
      W^J_{\nu}{}^{\rho}W^K_{\rho}{}^{\mu},
\end{aligned}
\]
and \(O_{HX}=(H^\dagger H)X_{\mu\nu}X^{\mu\nu}\),
\(O_{H\widetilde X}=(H^\dagger H)\widetilde X_{\mu\nu}X^{\mu\nu}\)
for \(X=G,W,B\), while
\[
  O_{HWB}=(H^\dagger\tau^I H)W^I_{\mu\nu}B^{\mu\nu},
  \qquad
  O_{H\widetilde WB}=(H^\dagger\tau^I H)\widetilde W^I_{\mu\nu}B^{\mu\nu}.
\]

The two-fermion part contains nineteen operator types:
\[
\begin{aligned}
\psi^2H^3&:\quad
  O_{eH},O_{uH},O_{dH},\\
\psi^2H^2D&:\quad
  O_{H\ell}^{(1)},O_{H\ell}^{(3)},
  O_{Hq}^{(1)},O_{Hq}^{(3)},
  O_{He},O_{Hu},O_{Hd},O_{Hud},\\
\psi^2XH&:\quad
  O_{eB},O_{eW},
  O_{uG},O_{uW},O_{uB},
  O_{dG},O_{dW},O_{dB}.
\end{aligned}
\]
The first line means the Yukawa monomials of
Definition~\ref{def:sm-higgs-yukawa-sector} multiplied by
\((H^\dagger H)\).  The \(H^2D\) line is generated by the templates
\[
\begin{gathered}
  (H^\dagger i\overleftrightarrow D_\mu H)
  (\bar\psi\gamma^\mu\psi),
  \qquad
  (H^\dagger i\overleftrightarrow D_\mu^I H)
  (\bar\psi\tau^I\gamma^\mu\psi),\\
  (\widetilde H^\dagger iD_\mu H)(\bar u\gamma^\mu d),
\end{gathered}
\]
with the displayed labels specifying the allowed SM representations.  The
dipole line means \((\bar\psi\sigma^{\mu\nu}\psi')HX_{\mu\nu}\), with \(H\)
or \(\widetilde H\) chosen so that the monomial has the same gauge quantum
numbers as the corresponding Yukawa monomial.

The baryon-number preserving four-fermion part contains twenty-five
one-generation operator types:
\[
\begin{aligned}
LL&:\quad
  O_{\ell\ell},O_{qq}^{(1)},O_{qq}^{(3)},
  O_{\ell q}^{(1)},O_{\ell q}^{(3)},\\
RR&:\quad
  O_{ee},O_{uu},O_{dd},O_{eu},O_{ed},
  O_{ud}^{(1)},O_{ud}^{(8)},\\
LR_{\rm vector}&:\quad
  O_{\ell e},O_{\ell u},O_{\ell d},O_{qe},
  O_{qu}^{(1)},O_{qu}^{(8)},O_{qd}^{(1)},O_{qd}^{(8)},\\
LR_{\rm scalar/tensor}&:\quad
  O_{\ell edq},O_{\ell equ}^{(1)},O_{\ell equ}^{(3)},
  O_{quqd}^{(1)},O_{quqd}^{(8)}.
\end{aligned}
\]
Flavor labels \(I,J,K,L\) are part of the coefficient coordinates.  When
flavor labels are restored, Hermiticity and exchange symmetries must be fixed
before counting real parameters.  Thus the basis is an ordered coordinate
system with specified flavor and reality conventions.
\end{definition}

\begin{definition}[Baryon-violating dimension-six classes]
\label{def:sm-baryon-violating-dimension-six-classes}
If baryon number is not imposed as a restriction on the effective couplings,
four additional gauge-invariant dimension-six classes are allowed.  In
standard chiral four-component notation they may be represented as
\[
\begin{aligned}
  O_{duq\ell}&=
  \epsilon_{abc}\epsilon_{ij}
  (d^a C u^b)(q^{ci}C\ell^j),\\
  O_{qque}&=
  \epsilon_{abc}\epsilon_{ij}
  (q^{ai}Cq^{bj})(u^cCe),\\
  O_{qqq\ell}&=
  \epsilon_{abc}\epsilon_{i\ell}\epsilon_{jk}
  (q^{ai}Cq^{bj})(q^{ck}C\ell^\ell),\\
  O_{duue}&=
  \epsilon_{abc}(d^aCu^b)(u^cCe).
\end{aligned}
\]
Each class carries flavor labels and changes \(B\) and \(L\) by one unit
with \(\Delta(B-L)=0\).  In the minimal SM definition used here these
coefficients are set to zero by imposing baryon number on the effective
couplings; in a more general effective theory they are independent
coordinates constrained by proton-decay observables.
\end{definition}

\begin{proposition}[Field-content exhaustion at dimension six]
\label{prop:sm-dimension-six-field-content-exhaustion}
Assume the field dimensions
\[
  \dim H=1,\qquad \dim D=1,\qquad \dim X_{\mu\nu}=2,\qquad
  \dim\psi=\frac32 .
\]
Every Lorentz-invariant local monomial of total engineering dimension six,
after integration by parts and algebraic identities but before imposing the
equations of motion, has one of the following field contents:
\[
  X^3,\quad H^6,\quad H^4D^2,\quad X^2H^2,\quad
  \psi^2H^3,\quad \psi^2H^2D,\quad \psi^2XH,\quad \psi^4 .
\]
Consequently Definitions~\ref{def:sm-warsaw-type-dimension-six-basis} and
\ref{def:sm-baryon-violating-dimension-six-classes} are exhaustive as a
dimension-six field-content classification.
\end{proposition}

\begin{proof}
A Lorentz scalar must contain an even number of spinor fields, so the number
of fermions is \(0,2\), or \(4\); six fermions already have dimension nine.
With no fermions, nonnegative integer solutions of
\[
  n_H+n_D+2n_X=6
\]
that cannot be reduced by integration by parts or by commutators of
covariant derivatives are represented by \(H^6\), \(H^4D^2\), \(X^2H^2\),
and \(X^3\).  Terms with one field strength and only Higgs fields have
uncontracted Lorentz indices unless derivatives are present; after commuting
derivatives, their field-strength part is absorbed into the displayed
classes or into the equation-of-motion quotient.

With two fermions, the remaining engineering dimension is three.  The only
possibilities are \(H^3\), \(H^2D\), and \(XH\), giving
\(\psi^2H^3\), \(\psi^2H^2D\), and \(\psi^2XH\).  With four fermions, the
fermions already have dimension six, so no additional \(H,D\), or \(X\) may
appear.  Gauge invariance and the finite Fierz identities then select the
contractions listed in the basis definitions.  The latter selection is a
finite algebraic basis choice inside \(\mathcal Q_6\), while this proposition
establishes that no further field-content class exists.
\end{proof}

\section{QCD Theta Angle and Quark-Mass Phase}

\begin{proposition}[Physical strong-CP phase coordinate]
\label{prop:sm-strong-cp-phase}
In the quark sector with nonsingular mass matrices, the invariant phase
coordinate involving the QCD theta angle and the quark mass matrices is
\[
  \overline\theta
  =
  \theta_3+\arg\det(M_uM_d)
  \quad\text{mod }2\pi .
  \label{eq:sm-strong-cp-phase}
\]
Equivalently, replacing \(M_u,M_d\) by positive diagonal mass matrices by
chiral flavor-coordinate changes shifts the coefficient of
\(\operatorname{tr}(F\wedge F)\) by the opposite amount, leaving
\(\overline\theta\) unchanged.
\end{proposition}

\begin{proof}
The determinant phase of the quark mass term changes under anomalous chiral
flavor rotations.  For a single Dirac quark of mass \(m\ee^{\ii\varphi}\), an
axial rotation by angle \(-\varphi/2\) makes the mass real.  The same
rotation changes the fermion measure by the four-dimensional chiral anomaly
and shifts the QCD theta angle by \(+\varphi\) in the convention of
Chapter~\ref{ch:volume-ii-20}.  For several quarks the phases
add, because the anomalous Jacobian is additive over left-handed Weyl
components in the fundamental color representation.  The total phase removed
from the quark mass matrices is \(\arg\det(M_uM_d)\).  Therefore the
combination \(\theta_3+\arg\det(M_uM_d)\) is invariant under the
mass-diagonalizing chiral coordinate choice.  If one quark mass matrix has a
zero singular value, the determinant phase is not a coordinate on the mass
chart and the theta dependence must be reconsidered separately.
\end{proof}

The empirical bound on \(\overline\theta\) is a bound on the coefficient of a
specific CP-odd QCD operator in the Standard Model definition.  It is not a
principle for choosing parameters.

\section{One-Loop Running in a Minimal Coupling Chart}

Renormalization-group formulae in the Standard Model are statements about a
renormalized coordinate system for couplings and masses.  The equations below
use dimensional regularization and minimal subtraction in a mass-independent
unbroken-representation chart with three active generations and one Higgs
doublet.  Threshold matching, the broken-phase expansion, and
higher-dimensional Wilson coefficients in an EFT extension are separate data.
The explicit-coupling convention is
the one fixed in Convention~\ref{conv:sm-electroweak-trace-normalization}.
Write
\[
  t:=\log\mu,\qquad \beta_x:=\frac{\dd x}{\dd t}.
\]
The hypercharge coupling \(g_1\) is the coupling to \(Y\), not the
grand-unified rescaling of it.

\begin{proposition}[One-loop gauge beta coefficients]
\label{prop:sm-one-loop-gauge-beta}
For three generations and one Higgs doublet, the one-loop gauge beta
functions in the convention above are
\[
  (4\pi)^2\beta_{g_1}
  =
  \frac{41}{6}g_1^3,\qquad
  (4\pi)^2\beta_{g_2}
  =
  -\frac{19}{6}g_2^3,\qquad
  (4\pi)^2\beta_{g_3}
  =
  -7g_3^3 .
  \label{eq:sm-one-loop-gauge-beta}
\]
\end{proposition}

\begin{proof}
Use the one-loop coefficient in a gauge theory with left-handed Weyl
fermions and complex scalars,
\[
  b_G=-\frac{11}{3}C_2(G)
      +\frac{2}{3}\sum_{\text{Weyl }f}T_G(R_f)
      +\frac{1}{3}\sum_{\text{complex }s}T_G(R_s),
  \qquad
  (4\pi)^2\beta_g=b_Gg^3 .
  \label{eq:sm-one-loop-general-gauge-beta}
\]
This is the background-field one-loop pole coefficient derived in
Chapter~\ref{ch:volume-ii-19}; here we only evaluate the representation
sums in the Standard Model coupling chart.  For \(U(1)_Y\) the same formula is
read with \(T_G(R)\) replaced by \(Y^2\), including all nonabelian spectator
multiplicities.  One generation contributes
\[
\begin{aligned}
  \sum_{\rm Weyl}Y^2
  &=
  6\left(\frac{1}{6}\right)^2
  +3\left(-\frac{2}{3}\right)^2
  +3\left(\frac{1}{3}\right)^2
  +2\left(-\frac{1}{2}\right)^2
  +1^2\\
  &=\frac{10}{3}.
\end{aligned}
\]
Three generations give \(10\), and the Higgs doublet contributes
\[
  \sum_{\rm scalar}Y^2=2\left(\frac{1}{2}\right)^2=\frac{1}{2}.
\]
Thus
\[
  b_1=\frac{2}{3}\cdot10+\frac{1}{3}\cdot\frac{1}{2}
  =\frac{41}{6}.
\]

For \(SU(2)_L\), \(C_2(SU(2))=2\) and \(T(\mathbf2)=1/2\) in the conventional
half-trace normalization used for the explicit coupling \(g_2\).  One
generation has three quark doublets and one lepton doublet, hence
\[
  \sum_{\rm Weyl}T_{SU(2)}=3\cdot\frac{1}{2}+\frac{1}{2}=2 .
\]
For three generations this is \(6\), and the Higgs doublet contributes
\(1/2\).  Therefore
\[
  b_2=-\frac{11}{3}\cdot2+\frac{2}{3}\cdot6+\frac{1}{3}\cdot\frac{1}{2}
  =-\frac{19}{6}.
\]

For \(SU(3)_c\), \(C_2(SU(3))=3\) and \(T(\mathbf3)=T(\overline{\mathbf3})=1/2\).
One generation contributes
\[
  2\cdot\frac{1}{2}+\frac{1}{2}+\frac{1}{2}=2,
\]
where the factor \(2\) is the weak multiplicity of \(Q_L\).  Three
generations give \(6\), and there is no colored scalar in the minimal field
content.
Thus
\[
  b_3=-\frac{11}{3}\cdot3+\frac{2}{3}\cdot6=-7 .
\]
\end{proof}

\begin{remark}[Hypercharge normalization]
If \(g_1^{\rm GUT}:=\sqrt{5/3}\,g_1\), then
\[
  (4\pi)^2\beta_{g_1^{\rm GUT}}
  =
  \frac{41}{10}(g_1^{\rm GUT})^3 .
\]
This is only a coordinate change.  The electroweak mass formulae in this
chapter use the unrescaled \(g_1\).
\end{remark}

\begin{quotedtheorem}[One-loop top-Higgs subsystem]
\label{qthm:sm-one-loop-top-higgs-subsystem}
Restrict the minimal Standard Model coupling chart to the invariant subspace in
which the only nonzero Yukawa singular value is the top coordinate \(y_t\).
In the same minimal-subtraction convention,
\[
  (4\pi)^2\beta_{y_t}
  =
  y_t\left[
    \frac{9}{2}y_t^2
    -8g_3^2
    -\frac{9}{4}g_2^2
    -\frac{17}{12}g_1^2
  \right],
  \label{eq:sm-one-loop-top-yukawa-beta}
\]
and
\[
\begin{aligned}
  (4\pi)^2\beta_{\lambda}
  &=
  24\lambda^2
  +12\lambda y_t^2
  -6y_t^4
  -3\lambda(3g_2^2+g_1^2)\\
  &\quad
  +\frac{3}{8}\left[
    2g_2^4+(g_2^2+g_1^2)^2
  \right].
\end{aligned}
  \label{eq:sm-one-loop-lambda-beta}
\]
\end{quotedtheorem}

\noindent
The restriction to one Yukawa singular value is a coordinate restriction, not
an estimate: the one-loop matrix beta functions are polynomial in the Yukawa
matrices and preserve the subspace where \(Y_d=Y_e=0\) and \(Y_u\) has only
one nonzero singular value.  The displayed equations are the corresponding
one-loop minimal-subtraction pole coefficients for the \(Q_3Hu^c_3\) Yukawa
vertex and the four-Higgs 1PI vertex.  Their group factors are fixed as
follows.

For the top Yukawa term, the \(SU(3)\) gauge correction sees the fundamental
quadratic Casimir \(C_F=4/3\) on each of the two colored fermion legs and
gives \(-6C_Fg_3^2=-8g_3^2\).  The \(SU(2)\) correction sees the doublet
Casimir \(3/4\) on the \(Q_3\) and \(H\) legs in the Yukawa vertex and gives
the coefficient \(-9g_2^2/4\).  The hypercharge coefficient is the sum of the
corresponding vertex and wave-function pole factors in the convention
\(Q=T^3+Y\), giving
\[
  -\frac{17}{12}g_1^2 .
\]
The positive self-Yukawa coefficient \(9y_t^2/2\) is the sum of the
color-multiplicity vertex pole and the three external-field pole
renormalizations on this one-singular-value subspace.

For the scalar quartic, the \(24\lambda^2\) term is the one-loop four-Higgs
fish contribution in the normalization
\(V=\lambda(H^\dagger H-v^2/2)^2\).  The top loop gives \(-6y_t^4\), with
the minus sign from the closed fermion loop and the factor \(3\) from color;
the mixed scalar-Yukawa wave-function and vertex terms give
\(12\lambda y_t^2\).  Gauge corrections linear in \(\lambda\) multiply the
Higgs doublet quadratic Casimirs and hypercharge to give
\(-3\lambda(3g_2^2+g_1^2)\).  The pure gauge box and exchange diagrams give
the positive quartic polynomial
\[
  \frac{3}{8}\left[2g_2^4+(g_2^2+g_1^2)^2\right].
\]
The statement is an equation for renormalized perturbative coordinates.  It
does not assert a nonperturbative electroweak continuum limit, and it does
not turn the running of \(\lambda\) into a principle for choosing ultraviolet
boundary data.  It also is not the full physical Standard Model trajectory
unless the remaining Yukawa singular values and all higher-dimensional
coordinates are consistently set to zero in the chosen chart; its role here
is to isolate the leading top-Higgs mechanism in the quartic running.

\section{Higgs Quartic Running and Large-Field Claims}

The running quartic \(\lambda(\mu)\) is a coordinate in a chosen subtraction
scheme.  The sign of this coordinate is a useful perturbative fact, but it is
not by itself an observable and it is not a theorem about a nonperturbatively
constructed Standard Model.  A statement often called ``Higgs vacuum
stability'' therefore has to say which mathematical object it concerns.  It
may concern a minimal-subtraction trajectory, a perturbative effective
potential, a semiclassical decay functional, or an additional Wilsonian EFT
extension.  These are different objects.  The large-field effective-potential
analysis is meaningful only after the subtraction scheme, gauge convention,
field range, perturbative error estimate, omitted higher-dimensional
operators if an EFT extension is being used, and the relation between a
background Higgs coordinate and the observable or decay problem have been
specified.

\begin{definition}[High-scale RG continuation]
\label{def:sm-high-scale-rg-continuation}
Fix coupling coordinates \(c^A(\mu)\).  In the minimal perturbative electroweak
chart these include
\[
  g_1(\mu),\qquad g_2(\mu),\qquad g_3(\mu),\qquad
  y_t(\mu),\qquad \lambda(\mu),\qquad \ldots .
\]
Also fix a subtraction scheme, a loop order, and matching data
\(c^A(\mu_0)\) at a reference scale.  A high-scale RG continuation to
\(\mu_1>\mu_0\) is the solution of the specified ordinary differential
equation
\[
  \frac{\dd c^A}{\dd\log\mu}=\beta^A(c)
  \label{eq:sm-high-scale-rg-continuation-ode}
\]
on an interval where the perturbative chart and its error estimate are
specified.  This continuation is not the assertion that the Standard Model has
a physical cutoff \(\mu_1\).  It is a continuation of renormalized coordinates
inside a perturbative scheme.  To reinterpret \(\mu_1\) as a Wilsonian cutoff,
one must change the problem by adding a Wilsonian QFT definition at scale
\(\mu_1\), a matching map from that definition to the low-energy coordinates,
and a proof or controlled hypothesis that the observables under discussion are
reproduced in the overlap regime.
\end{definition}

\begin{remark}[Scope of minimal-subtraction running]
\label{rem:sm-ms-running-and-wilsonian-data}
The minimal-subtraction beta functions are obtained from the pole parts of
renormalized 1PI coordinate changes in dimensional regularization.  Solving
\eqref{eq:sm-high-scale-rg-continuation-ode} transports coordinates in this
renormalized coupling chart.  A Wilsonian cutoff theory is different data: it
requires a finite regulator or BV pushforward, a cutoff-dependent local
action or half-density, and a statement that low-momentum observables are
preserved by integrating out degrees of freedom.  The ordinary differential
equation for minimal-subtraction coordinates supplies only the coordinate
transport.  Treating the endpoint of that trajectory as a physical
high-cutoff Wilsonian EFT is an additional assumption about the theory under
discussion, and the assumption must be stated with its regulator or BV data
and its matching map.
\end{remark}

\begin{remark}[Extensions beyond the Standard Model]
A proposed extension of the Standard Model is an additional mathematical
hypothesis, not an explanation by itself.  If it is used only through finitely
many higher-dimensional operators, it is a choice of EFT coordinates together
with matching assumptions.  If it is claimed to be an ultraviolet completion,
the claim requires a regulator or constructive framework at least as explicit
as the ones demanded in
Definition~\ref{def:sm-finite-chiral-gauge-regulator-data} and
Remark~\ref{rem:sm-ms-running-and-wilsonian-data}.  A perturbative
Lagrangian, a high-scale RG trajectory, or a phenomenological motivation is
not by itself such a theory.
\end{remark}

\begin{definition}[Quartic positivity in a specified chart]
\label{def:sm-quartic-positivity-statement}
Fix a subtraction scheme, a gauge for the perturbative effective potential if
one is used, a loop order, matching maps from measured pole or threshold
observables to coupling coordinates at \(\mu_0\), and a scale interval
\(I=[\mu_0,\mu_1]\).  Specify whether \(I\) is only a high-scale RG
continuation in the sense of
Definition~\ref{def:sm-high-scale-rg-continuation}, or whether an
additional Wilsonian or constructive definition is being attached to the
endpoint.
A quartic-positivity statement on \(I\) is the claim that the solution of the
specified RG equations satisfies
\[
  \lambda(\mu)>0,\qquad \mu\in I,
  \label{eq:sm-quartic-positive-trajectory}
\]
with an error estimate that includes threshold matching, loop truncation,
input-parameter uncertainties, and the effect of higher-dimensional
operators in the chosen coupling chart.  It is a statement about
perturbative coordinates.  It is not a nonperturbative vacuum statement
unless the additional regulator and observable data below make it one.
\end{definition}

\begin{definition}[Metastability in a specified semiclassical calculation]
\label{def:sm-metastability-statement}
In the same chart, a metastability claim is not the assertion
\(\lambda(\mu)<0\).  It is a semiclassical statement about a
false-vacuum decay functional: one must specify the gauge-fixed effective
action used for the bounce calculation, the fluctuation determinant
convention, the renormalization scale-setting prescription, the spacetime
volume or cosmological history against which the rate is compared, and the
higher-dimensional operators retained in the field range of the bounce.  The
output is an inequality of the form
\[
  \Gamma_{\rm decay}\,{\mathcal V}_4 < \varepsilon ,
  \label{eq:sm-metastability-rate-statement}
\]
with a specified uncertainty.  The sign of \(\lambda\) along an RG trajectory
is one input to such a calculation, not the calculation itself.  It becomes a
prediction of a QFT only after the path integral, field range, and omitted
operator remainder are controlled in the regime sampled by the bounce.
\end{definition}

\begin{proposition}[Large-field EFT expansion requires a field-domain
hypothesis]
\label{prop:sm-radial-potential-dimension-six}
Let \(h\) be the radial Higgs coordinate defined by
\[
  H^\dagger H=\frac{h^2}{2}.
\]
In a coupling chart whose scalar potential is
\[
  V_{\leq 6}(h)
  =
  \lambda\left(\frac{h^2-v^2}{2}\right)^2
  +
  \frac{C_H}{\Lambda^2}\left(\frac{h^2}{2}\right)^3
  \label{eq:sm-higgs-radial-potential-dim-six}
\]
the large-field quartic approximation is controlled only for
\(h^2/\Lambda^2\) in the specified EFT range.  More precisely, for
\(h\gg v\),
\[
  V_{\leq 6}(h)
  =
  \frac{\lambda}{4}h^4
  \left[
    1+\frac{C_H}{2\lambda}\frac{h^2}{\Lambda^2}
    +O\!\left(\frac{v^2}{h^2}\right)
  \right],
  \label{eq:sm-higgs-dim-six-relative-size}
\]
whenever \(\lambda\neq0\).  Therefore a sign statement about the
dimension-four coordinate \(\lambda\) alone is not a statement about the
large-field potential unless the higher-dimensional coordinates and the field
range have been specified.  More generally, an operator
\[
  \frac{C_n}{\Lambda^{2n-4}}(H^\dagger H)^n,\qquad n\geq3,
  \label{eq:sm-higgs-general-large-field-operator}
\]
contributes, relative to the quartic term, by the factor
\[
  \frac{C_n}{\lambda\,2^{\,n-2}}
  \left(\frac{h^2}{\Lambda^2}\right)^{n-2}.
  \label{eq:sm-higgs-general-large-field-ratio}
\]
Thus a finite truncation of the large-field potential is a controlled EFT
statement only in a field domain where the omitted tower is bounded by a
specified remainder.
\end{proposition}

\begin{proof}
Substituting \(H^\dagger H=h^2/2\) gives
\[
  \lambda\left(\frac{h^2-v^2}{2}\right)^2
  =
  \frac{\lambda}{4}h^4
  -\frac{\lambda v^2}{2}h^2
  +\frac{\lambda v^4}{4}
\]
and
\[
  \frac{C_H}{\Lambda^2}\left(\frac{h^2}{2}\right)^3
  =
  \frac{C_H}{8\Lambda^2}h^6 .
\]
Factoring \(\lambda h^4/4\) from the large-field expression gives
\eqref{eq:sm-higgs-dim-six-relative-size}.  The ratio of the dimension-six
term to the quartic term grows as \(h^2/\Lambda^2\).  Thus the EFT coupling
chart, not only the dimension-four coordinate, is part of the statement.

For the general operator \eqref{eq:sm-higgs-general-large-field-operator},
substitution of \(H^\dagger H=h^2/2\) gives
\[
  \frac{C_n}{\Lambda^{2n-4}}\frac{h^{2n}}{2^n}.
\]
Dividing by \(\lambda h^4/4\) gives
\eqref{eq:sm-higgs-general-large-field-ratio}.  If \(h^2/\Lambda^2\) is not
small, or if the coefficients \(C_n\) are not bounded by a stated matching
hypothesis, no finite truncation gives a controlled statement about the
potential in that region.
\end{proof}

\begin{remark}[Large-field claims in the hybrid definition]
\label{rem:sm-large-field-claims-hybrid-definition}
The definition contains the electroweak sector as a renormalized perturbative
chiral gauge theory and supplies matching maps to lower-energy EFTs and
nonperturbative QCD inputs.  A nonperturbative large-field stability or
metastability theorem for the Higgs coordinate would require further data:
a probability measure, Hilbert space, or operator algebra for the full chiral
electroweak theory, or a Wilsonian/EFT construction with a declared field
range and matching map.  Minimal-subtraction coordinate transport has the
scope described in Remark~\ref{rem:sm-ms-running-and-wilsonian-data}.  If one
changes the problem to an EFT extension with higher-dimensional operators,
Proposition~\ref{prop:sm-radial-potential-dimension-six} shows that a
large-field statement also samples the omitted operator coordinates unless the
field range and matching assumptions control them.  The baseline hybrid
definition therefore supports conditional perturbative statements about a
chosen coupling chart or decay functional; an unconditional QFT theorem about
Higgs large-field stability or metastability is a separate construction
problem.
\end{remark}

\begin{remark}[One-loop source of downward quartic running]
\label{rem:sm-top-source-downward-quartic}
In the one-loop top-Higgs subsystem of
Quoted Theorem~\ref{qthm:sm-one-loop-top-higgs-subsystem}, the beta function at
\(\lambda=0\) is
\[
  (4\pi)^2\beta_\lambda\big|_{\lambda=0}
  =
  -6y_t^4
  +
  \frac{3}{8}\left[
    2g_2^4+(g_2^2+g_1^2)^2
  \right].
  \label{eq:sm-lambda-beta-at-zero}
\]
Thus the top-Yukawa term supplies the negative entry in this
minimal-subtraction quartic beta function.  The statement belongs to the
specified perturbative coupling chart.
\par

Set \(\lambda=0\) in \eqref{eq:sm-one-loop-lambda-beta}.  The terms
\(24\lambda^2\), \(12\lambda y_t^2\), and
\(-3\lambda(3g_2^2+g_1^2)\) vanish, leaving exactly
\eqref{eq:sm-lambda-beta-at-zero}.  The negative sign of the top term is the
closed-fermion-loop sign already identified in the proof of
Quoted Theorem~\ref{qthm:sm-one-loop-top-higgs-subsystem}.  No
scheme-independent observable is obtained until the matching inputs and the
observable or decay functional have been specified.
\end{remark}

\begin{definition}[Data required for a Higgs large-field claim]
\label{def:sm-higgs-large-field-claim-data}
A Higgs large-field claim in the hybrid Standard Model definition has a
precise meaning only after specifying:
\begin{enumerate}
\item the input observables and matching maps used to determine
      \(g_1,g_2,g_3,y_t,\lambda\) and the relevant higher-dimensional
      coefficients at the reference scale;
\item the subtraction scheme, loop order, threshold convention, and scale
      interval for the RG evolution;
\item whether the endpoint of the scale interval is only a formal RG
      continuation of renormalized coordinates or is being supplemented by an
      additional Wilsonian, lattice, BV, or constructive definition;
\item whether the output is a quartic-positivity statement in the sense of
      Definition~\ref{def:sm-quartic-positivity-statement} or a metastability
      statement in the sense of
      Definition~\ref{def:sm-metastability-statement};
\item the treatment of \(O_H=(H^\dagger H)^3\) and every other retained
      higher-dimensional operator whose contribution is not uniformly small
      over the field range;
\item a remainder estimate for perturbative truncation and for omitted
      operators.
\end{enumerate}
Without these data, a sentence about ``vacuum stability'' has no definite
mathematical content in this chapter.  Even with these data, the result is
conditional unless a theorem or controlled argument relates the chosen
perturbative quantity to the claimed observable or decay problem.
\end{definition}

\section{Precision Electroweak Self-Energy Coordinates}

Precision electroweak fits use a finite-dimensional projection of the
renormalized electroweak two-point functions.  The projection is meaningful
only after the input scheme, subtraction scheme, and class of observables have
been fixed.  This section defines the \(S,T,U\) coordinates as such a
two-point projection.

\begin{definition}[Transverse self-energy difference]
\label{def:sm-transverse-self-energy-difference}
Fix a reference electroweak coupling chart with the same input values of
\(\alpha_{\rm em}\), \(G_F\), and \(m_Z\).  Let
\[
  s_W:=\frac{g_1}{\sqrt{g_1^2+g_2^2}},
  \qquad
  c_W:=\frac{g_2}{\sqrt{g_1^2+g_2^2}},
  \qquad
  e:=g_2s_W=g_1c_W ,
\]
so that \(m_W^2=c_W^2m_Z^2\) at tree level in the chart of
Proposition~\ref{prop:sm-electroweak-stabilizer-masses}.  For
\(V,V'\in\{W,Z,\gamma\}\), define \(\Delta\Pi_{VV'}(q^2)\) to be the
difference between the renormalized transverse 1PI two-point scalar in the
chart under study and the corresponding scalar in the reference chart,
with identical external-current normalization.  Thus
\[
  \Delta\Gamma^{\mu\nu}_{VV'}(q)\big|_{\rm trans}
  =
  -\ii
  \left(\eta^{\mu\nu}-\frac{q^\mu q^\nu}{q^2}\right)
  \Delta\Pi_{VV'}(q^2),
  \label{eq:sm-transverse-self-energy-definition}
\]
on conserved external currents.  The electromagnetic on-shell conditions are
chosen so that
\[
  \Delta\Pi_{\gamma\gamma}(0)=0,
  \qquad
  \Delta\Pi_{Z\gamma}(0)=0 .
  \label{eq:sm-oblique-em-onshell}
\]
Assume, for the definition below, that the self-energy differences are
analytic at \(q^2=0\) on the relevant sheet and write
\(\Delta\Pi'_{VV'}(0):=\dd\Delta\Pi_{VV'}(q^2)/\dd q^2|_{q^2=0}\).
\end{definition}

\begin{definition}[\(S,T,U\) two-point coordinates]
\label{def:sm-stu-coordinates}
Under the hypotheses of
Definition~\ref{def:sm-transverse-self-energy-difference}, the
Peskin--Takeuchi \(S,T,U\) coordinates are the three finite combinations
\[
\begin{aligned}
  \alpha_{\rm em} S
  &=
  4s_W^2c_W^2
  \left[
    \Delta\Pi'_{ZZ}(0)
    -\frac{c_W^2-s_W^2}{s_Wc_W}\Delta\Pi'_{Z\gamma}(0)
    -\Delta\Pi'_{\gamma\gamma}(0)
  \right],\\
  \alpha_{\rm em} T
  &=
  \frac{\Delta\Pi_{WW}(0)}{m_W^2}
  -
  \frac{\Delta\Pi_{ZZ}(0)}{m_Z^2},\\
  \alpha_{\rm em} U
  &=
  4s_W^2
  \left[
    \Delta\Pi'_{WW}(0)
    -c_W^2\Delta\Pi'_{ZZ}(0)
    -2s_Wc_W\Delta\Pi'_{Z\gamma}(0)
    -s_W^2\Delta\Pi'_{\gamma\gamma}(0)
  \right].
\end{aligned}
\label{eq:sm-stu-definitions}
\]
Here
\(\alpha_{\rm em}:=e^2/(4\pi)\).  These are coordinates on the two-point part
of the electroweak coupling chart, not on the full space of all possible
four-fermion, vertex, and detector corrections.
\end{definition}

\begin{remark}[What the \(T\) coordinate measures]
\label{rem:sm-t-coordinate-rho}
Suppose the only low-momentum effect in the two-point chart is a shift of the
charged and neutral weak mass coordinates,
\[
  m_W^2\mapsto m_W^2+\delta m_W^2,
  \qquad
  m_Z^2\mapsto m_Z^2+\delta m_Z^2,
\]
with \(s_W,c_W\) held fixed by the chosen input convention and with
\(\Delta\Pi_{WW}(0)=\delta m_W^2\),
\(\Delta\Pi_{ZZ}(0)=\delta m_Z^2\).  Then
\[
  \alpha_{\rm em}T
  =
  \frac{\delta m_W^2}{m_W^2}
  -
  \frac{\delta m_Z^2}{m_Z^2}
  \label{eq:sm-t-fractional-mass-shift}
\]
and the first-order shift of
\[
  \rho:=\frac{m_W^2}{m_Z^2c_W^2}
\]
is
\[
  \delta\rho=\alpha_{\rm em}T+O(\delta^2).
  \label{eq:sm-t-rho-shift}
\]
The first equality is exactly the definition of \(T\) after the stated
identification of the zero-momentum self-energies with mass-coordinate shifts.
For the second, use the tree identity \(m_W^2=m_Z^2c_W^2\) and expand
\[
  \rho_{\rm shifted}
  =
  \frac{m_W^2+\delta m_W^2}
       {(m_Z^2+\delta m_Z^2)c_W^2}
  =
  \left(1+\frac{\delta m_W^2}{m_W^2}\right)
  \left(1+\frac{\delta m_Z^2}{m_Z^2}\right)^{-1}.
\]
The first-order Taylor expansion of \((1+x)^{-1}\) gives
\[
  \rho_{\rm shifted}
  =
  1+
  \frac{\delta m_W^2}{m_W^2}
  -
  \frac{\delta m_Z^2}{m_Z^2}
  +O(\delta^2),
\]
which proves \eqref{eq:sm-t-rho-shift}.  This is the precise sense in which
\(T\) records custodial-symmetry breaking in a two-point electroweak chart.
\end{remark}

\begin{proposition}[Domain of the \(S,T,U\) projection]
\label{prop:sm-stu-domain}
Assume all corrections to the relevant electroweak observables enter through
the analytic transverse two-point functions in
Definition~\ref{def:sm-transverse-self-energy-difference}, and keep only the
constant and first-derivative Taylor coefficients at \(q^2=0\).  After the
input coordinates \(\alpha_{\rm em}\), \(G_F\), and \(m_Z\) are fixed, the
remaining three input-independent two-point combinations can be chosen as the
coordinates \(S,T,U\) in Definition~\ref{def:sm-stu-coordinates}.
\end{proposition}

\begin{proof}
The tree exchange amplitudes of conserved external fermion currents are
obtained by inverting the quadratic form of the gauge fields and contracting
with the current vertices.  To first order in
\(\Delta\Pi\), the correction is
\[
  D_0(q)\,\Delta\Pi(q^2)\,D_0(q),
\]
where \(D_0\) is the tree propagator matrix in the same current basis.  Under
the analyticity hypothesis, the correction through order \(q^2\) depends only
on \(\Delta\Pi_{VV'}(0)\) and \(\Delta\Pi'_{VV'}(0)\).  Electromagnetic gauge
invariance and the on-shell electromagnetic normalization impose
\eqref{eq:sm-oblique-em-onshell}.  Finite redefinitions of the three input
coordinates \(\alpha_{\rm em}\), \(G_F\), and \(m_Z\) remove three further
linear combinations: the electric charge normalization, the charged-current
normalization used to define \(G_F\), and the reference neutral-pole mass.

The combinations displayed in \eqref{eq:sm-stu-definitions} are a convenient
basis of the remaining three-dimensional quotient.  The \(T\) coordinate is
the residual custodial mass-ratio coordinate by
Remark~\ref{rem:sm-t-coordinate-rho}; the \(S\) coordinate is the
neutral-current derivative combination that remains after the photon charge
normalization is removed; and \(U\) is the corresponding charged-minus-neutral
derivative coordinate.  If nonanalytic thresholds, nonlocal detector
functionals, vertex corrections, box corrections, or independent local
four-fermion operators are kept at the same order, they are additional
coordinates of the matched prediction and are not represented by \(S,T,U\)
alone.
\end{proof}

\section{The Hybrid Definition}

\begin{definition}[QCD--electroweak coupling data]
\label{def:sm-qcd-ew-coupling-data}
Let \(\mathfrak D_{\rm QCD}^{\rm W}\) be a Wightman QFT with Hilbert space
\(\mathcal H_{\rm QCD}\), vacuum \(\Omega_{\rm QCD}\), and a net or common
domain of color-gauge-invariant local operator-valued distributions.  A
QCD--electroweak coupling data \(\mathcal C_{\rm QCD\text{-}EW}\) consist
of the following items.
\begin{enumerate}
\item A finite or countable list of renormalized color-gauge-invariant local
      QCD fields
      \[
        \mathcal O_A(x)\in\mathcal O_{\rm QCD}^{\rm inv},
      \]
      including the flavor vector and axial currents, charged weak currents,
      electromagnetic current, and scalar or pseudoscalar quark-density
      operators needed after an electroweak scalar chart is chosen.  Each
      \(\mathcal O_A\) is an operator-valued tempered distribution, with its
      domain, normalization, and composite-operator scheme specified.
\item The transformation laws of these operators under the flavor and
      electroweak source symmetries that are to be gauged perturbatively,
      together with their Ward identities, anomaly contact terms, and
      current-conservation or partial-conservation equations.
\item A local perturbative interaction functional
      \[
        I_{\rm int}[\varphi_{\rm EW};\mathcal O]
        =
        \sum_A \int_{M_4}\varphi_{\rm EW}^A(x)\,\mathcal O_A(x),
        \label{eq:sm-qcd-ew-hybrid-coupling}
      \]
      where \(\varphi_{\rm EW}^A\) denotes the corresponding perturbative
      electroweak field, source, or composite expression.  The product in
      \eqref{eq:sm-qcd-ew-hybrid-coupling} is required to be invariant under
      the electroweak gauge symmetry after the perturbative gauge-fixing or BV
      formulation has been specified.
\item A prescription for mixed correlators as formal series in electroweak
      couplings whose coefficients are Wightman distributions of the QCD
      operators \(\mathcal O_A\), combined with the perturbative electroweak
      propagators and vertices.  The prescription includes the renormalization
      of coincident insertions and the contact-term scheme.
\end{enumerate}
\end{definition}

\begin{proposition}[Hybrid coupling is additional structure]
\label{prop:sm-hybrid-coupling-additional-structure}
The data \(\mathcal C_{\rm QCD\text{-}EW}\) in
Definition~\ref{def:sm-qcd-ew-coupling-data} is not determined by separately
giving a nonperturbative QCD Wightman theory and a perturbative electroweak
Lagrangian.  It is additional mathematical structure needed to define the
hybrid Standard Model prediction framework.
\end{proposition}

\begin{proof}
A nonperturbative QCD Wightman theory contains physical local fields only in
the color-gauge-invariant operator algebra.  The colored quark fields that
appear in a gauge-fixed perturbative Lagrangian are not, by themselves,
Wightman fields on the physical QCD Hilbert space.  Therefore an electroweak
boson or Higgs fluctuation cannot be coupled to a nonperturbative QCD sector
by reusing a colored perturbative quark line.  The coupling has to be
expressed through color-gauge-invariant QCD operator-valued distributions:
for example the electromagnetic current, charged weak currents, neutral
currents, and scalar density operators.  Their normalizations, contact terms,
Ward identities, and anomaly conventions determine the coefficients of mixed
correlators with electroweak external fields.

At order \(n\) in the electroweak couplings, a mixed correlator is a finite
sum of perturbative electroweak kernels paired with QCD Wightman
distributions of the form
\[
  \bigl\langle
    \Omega_{\rm QCD},
    \mathcal O_{A_1}(x_1)\cdots \mathcal O_{A_n}(x_n)
    \mathcal X_1(y_1)\cdots\mathcal X_m(y_m)
    \Omega_{\rm QCD}
  \bigr\rangle,
  \label{eq:sm-qcd-ew-mixed-coefficient}
\]
where the \(\mathcal X_i\) are the QCD operators appearing in the observable.
The distribution \eqref{eq:sm-qcd-ew-mixed-coefficient}, its extension to
coincident points, and its Ward identities are not supplied by the
electroweak perturbative Lagrangian.  Conversely, the pure QCD Wightman theory
does not know which electroweak source or gauge field multiplies which
operator.  Thus the hybrid coupling structure is logically necessary.
\end{proof}

\begin{definition}[Hybrid Standard Model definition]
\label{def:hybrid-standard-model-definition}
The Standard Model is used in this monograph through the following hybrid
definition, not as a single completed constructive theorem for a
four-dimensional continuum QFT:
\[
  \mathfrak D_{\rm SM}
  =
  \bigl(
    \mathfrak D_{\rm QCD}^{\rm W},
    \mathfrak D_{\rm EW}^{\rm pert},
    \mathcal C_{\rm QCD\text{-}EW},
    \mathfrak D_{\rm weak\,EFT},
    \mathcal M
  \bigr).
\]
Here \(\mathfrak D_{\rm QCD}^{\rm W}\) denotes the nonperturbative continuum
QCD Wightman theory selected by the \(SU(3)_c\) gauge theory with quark
matter; in constructive and computational practice it is approached through
finite lattice regulators and a continuum-limit problem of the kind stated in
Hypothesis~\ref{hyp:lattice-yang-mills-continuum-limit}.  When a concrete
lattice regulator is needed for a hadronic matrix element, it will be denoted
\(\mathfrak D_{\rm QCD}^{\rm lat}\); this notation refers to the regulator
used to approximate the QCD component, not to a cutoff definition of the full
Standard Model.
\(\mathfrak D_{\rm EW}^{\rm pert}\) denotes the renormalized electroweak
chiral gauge theory with Higgs field, constructed as a perturbative formal
power series to all orders after gauge fixing or BV treatment, anomaly
cancellation, and a subtraction scheme have been fixed;
\(\mathcal C_{\rm QCD\text{-}EW}\) is the hybrid coupling data of
Definition~\ref{def:sm-qcd-ew-coupling-data};
\(\mathfrak D_{\rm weak\,EFT}\) is the tower of lower-energy effective
theories obtained by integrating out massive weak fields and heavy flavors;
and \(\mathcal M\) is the matching data fixing parameters and local operators
between overlapping regimes.  The symbol
\(\mathfrak D_{\rm EW}^{\rm pert}\) is not a finite-cutoff definition and is
not a claim that a nonperturbative four-dimensional chiral electroweak
continuum construction has been supplied.
\end{definition}

\begin{definition}[Matching data in the hybrid Standard Model]
\label{def:sm-hybrid-matching-data}
The symbol \(\mathcal M\) in
Definition~\ref{def:hybrid-standard-model-definition} consists of the following
specified data for the observable class under consideration:
\begin{enumerate}
\item a sequence of scales or windows where adjacent descriptions overlap;
\item renormalized parameter charts for \(g_3,g_2,g_1,\lambda,v\), the
      Yukawa matrices, theta angles, and, when an EFT extension is being used,
      higher-dimensional Wilson coordinates;
\item threshold matching maps when \(W,Z,h,t,b,c,\tau,\ldots\) are removed
      from the active field content;
\item anomalous-dimension matrices for the chosen operator bases between
      thresholds;
\item the QCD--electroweak coupling data
      \(\mathcal C_{\rm QCD\text{-}EW}\), including the scheme for the
      currents, scalar densities, contact terms, and Ward identities entering
      the observable;
\item nonperturbative QCD matrix elements, defined by
      \(\mathfrak D_{\rm QCD}^{\rm lat}\) or by another explicitly specified
      nonperturbative regulator, for hadronic operators;
\item the error model or remainder statement attached to each perturbative,
      lattice, or combined extrapolation.
\end{enumerate}
A numerical precision claim is well defined only after these choices and
their error estimates are specified.
\end{definition}

\begin{remark}[Logical parts of the hybrid definition]
\label{rem:sm-hybrid-logical-layers}
Definition~\ref{def:hybrid-standard-model-definition} separates
nonperturbative QCD input, perturbative electroweak definitions, their hybrid
operator coupling, lower-energy effective theories, and observable-specific
matching.  These parts cannot be replaced by one another without adding an
extra theorem or hypothesis.
\par
First, the QCD sector is the nonperturbative part of the definition.  It is
intended as a continuum Wightman QFT with gauge-invariant local and extended
observables, while compact lattice link variables, Haar measure, and finite
Grassmann algebra for quarks provide the concrete nonperturbative regulators
used to define approximating matrix elements.  The existence of the
four-dimensional continuum limit with the desired physical spectrum is still a
theorem-level constructive problem, but the logical role of this component is
not perturbative electroweak renormalization.

Second, the electroweak sector is a perturbatively renormalizable chiral gauge
theory.  Its perturbative consistency uses anomaly cancellation, BRST or BV
Ward identities, and the local power-counting analysis of the BEH phase.  This
defines renormalized Green functions and operator insertions as formal power
series to all orders in the chosen subtraction scheme.  A standalone
nonperturbative continuum construction is not supplied by these facts.  The
hypercharge \(U(1)\) coupling is not asymptotically free, and the
four-dimensional scalar quartic sector lies on the same marginal-triviality
boundary discussed in Chapters~\ref{ch:scalar-path-integrals-green-functions}
and \ref{ch:volume-ii-ising-universality}.  Moreover, a fully satisfactory
interacting lattice construction of a four-dimensional chiral gauge theory
with the Standard Model representation content is a separate problem, related
to the determinant-line and chiral-lattice issues in
Chapter~\ref{ch:constructive-lattice-fermions-chiral-symmetry}.

Third, the coupling of these two parts is not automatic.  It is mediated by
the color-gauge-invariant QCD operator insertions specified in
\(\mathcal C_{\rm QCD\text{-}EW}\).  This is the point at which
nonperturbative hadronic matrix elements enter electroweak amplitudes and
where contact terms and Ward identities must be fixed.

Fourth, physical predictions are extracted in regimes where a controlled
object is available.  High-energy short-distance processes use perturbative
renormalization and factorization with explicitly stated infrared-safe
observables.  Low-energy hadronic weak processes use QCD matrix elements
defined by the QCD regulator together with weak effective operators matched at
the electroweak scale.  Precision electroweak observables use finite-order
renormalized perturbation theory.  A cutoff parameter is not part of the
prediction; any auxiliary regulator dependence has been removed or absorbed
into the stated renormalized coordinates and operator scheme.
\end{remark}

\section{Chiral Lattice Status of the Electroweak Sector}

The statement that the electroweak sector is a chiral gauge theory has a
precise regulator consequence.  Vectorlike lattice fermions define a finite
Grassmann integral by an ordinary determinant.  A chiral gauge theory instead
requires a gauge-covariant choice of a subspace of left-handed lattice
fermions, and the corresponding fermion integral is a section of a
determinant line over the finite gauge-field space.  The phase of that
section is part of the regulator, not an optional convention.

\begin{definition}[Finite chiral-gauge regulator data]
\label{def:sm-finite-chiral-gauge-regulator-data}
A finite chiral-gauge regulator for the Standard Model representation
content consists of the following data at lattice spacing \(a\) and finite
volume:
\begin{enumerate}
\item compact lattice gauge fields for the chosen global form of the local
      product data \(SU(3)_c\times SU(2)_L\times U(1)_Y\), with a local gauge
      action;
\item a local gauge-covariant lattice Dirac operator, for example an
      overlap/Ginsparg--Wilson operator satisfying
      \eqref{eq:lattice-ginsparg-wilson};
\item chiral projectors assigning \(Q_L,L_L\) to left-handed subspaces and
      \(u^c,d^c,e^c\) to the conjugate left-handed representations of
      Definition~\ref{def:sm-one-generation-left-weyl};
\item a gauge-invariant, local, and reflection-compatible choice of phase of
      the chiral fermion determinant line over lattice gauge fields;
\item lattice Higgs and Yukawa terms whose gauge quantum numbers reduce to
      Definition~\ref{def:sm-higgs-yukawa-sector} in the scaling window;
\item uniform locality and reflection-positivity, or a specified replacement
      reconstruction route, strong enough to define a candidate continuum
      local QFT.
\end{enumerate}
\end{definition}

\begin{proposition}[Local anomaly cancellation as determinant-line obstruction]
\label{prop:sm-chiral-lattice-anomaly-necessity}
For finite chiral-gauge regulator data in the sense of
Definition~\ref{def:sm-finite-chiral-gauge-regulator-data}, gauge
invariance of the chiral determinant line requires the perturbative gauge
anomaly coefficients of Proposition~\ref{prop:sm-generation-anomaly-cancellation}
to vanish.  It also requires the finite \(SU(2)\) doublet-parity condition in
that proposition.
\end{proposition}

\begin{proof}
At finite lattice spacing the Grassmann integral over a chiral subspace gives
a determinant of the chiral Dirac map between finite-dimensional spaces.  As
the background gauge field varies, these determinants form a complex line
bundle.  A gauge-invariant fermion measure is a nonvanishing section of this
line bundle whose phase is compatible with gauge transformations.  Around an
infinitesimal closed loop in gauge-field space, the phase holonomy is
computed by the curvature of the determinant connection.  In the classical
continuum limit, the local part of this curvature is the same descent
polynomial that appears in Chapter~\ref{ch:volume-ii-20}.  Therefore a
nonzero local gauge anomaly gives nontrivial infinitesimal holonomy along
gauge orbits and forbids a gauge-invariant phase.  The calculation of
Proposition~\ref{prop:sm-generation-anomaly-cancellation} is precisely the
vanishing of these local obstruction coefficients for the Standard Model
representations.

There is also a global obstruction for an odd number of \(SU(2)\) doublets:
transport around the nontrivial loop in the weak \(SU(2)\) gauge factor
changes the sign of the fermion Pfaffian.  The Standard Model has four weak
doublets per generation, hence avoids this obstruction.
\end{proof}

\begin{remark}[Scope of the anomaly-obstruction calculation]
The vanishing of the local curvature obstruction and the \(SU(2)\) parity
obstruction is necessary regulator data.  A complete finite chiral-gauge
regulator still has to construct a gauge-invariant local phase functional,
prove the required locality and reflection or reconstruction properties, and
control the four-dimensional continuum scaling limit.
\end{remark}

\begin{remark}[Lattice QCD inside the hybrid Standard Model]
\label{rem:sm-lattice-qcd-inside-hybrid-sm}
The lattice regulators used to approximate the QCD component
\(\mathfrak D_{\rm QCD}^{\rm W}\) in
Definition~\ref{def:hybrid-standard-model-definition} supply nonperturbative
definitions of gauge-invariant QCD matrix elements after the electroweak
matching data have selected the relevant color-singlet operators.  They do
so because lattice QCD uses the color gauge group and quark fields as a
vectorlike fermion system for the purpose of strong-interaction matrix
elements.  Its finite Grassmann integral is an ordinary determinant, or a
controlled product of determinants, coupled to compact \(SU(3)_c\) links.
The electroweak theory before matching has chiral assignments: \(Q_L\) is an
\(SU(2)_L\) doublet while \(u^c,d^c\) are weak singlets with different
hypercharges, and \(L_L,e^c\) have the analogous chiral electroweak
assignment.  A finite regulator for that system must solve the determinant
line and phase problem in
Definition~\ref{def:sm-finite-chiral-gauge-regulator-data}.

The hybrid construction avoids conflating these two uses.  For hadronic weak
observables, the weak fields are first matched onto gauge-invariant local
operators such as four-quark, semileptonic, or electromagnetic-current
operators.  The QCD lattice regulator then computes the matrix elements of
those color-singlet operators.  This gives a valid matched prediction after
the map \(\mathcal M\) is specified; a nonperturbative construction of the
microscopic chiral electroweak gauge theory is the distinct finite-regulator
problem described above.
\end{remark}

\begin{definition}[Standard Model prediction]
\label{def:sm-hybrid-prediction}
A Standard Model prediction for an observable \(O\), in the sense used in
this chapter, is a number or distribution obtained only after the following
data are specified:
\begin{enumerate}
\item the energy regime and the component of the hybrid definition used in
      that regime: nonperturbative QCD input, renormalized perturbative
      electroweak theory, a lower-energy EFT, or an additional Wilsonian/EFT
      extension;
\item the gauge-invariant local operator, infrared-safe inclusive observable,
      or detector functional representing \(O\);
\item the matching map from the adjacent higher-energy description, including
      the scheme and normalization of all operator coordinates;
\item the perturbative, lattice, or combined expansion parameter and the
      associated remainder or systematic-error class;
\item the positivity, unitarity, gauge-invariance, and infrared-safety
      statements needed for the observable.
\end{enumerate}
\end{definition}

\section{Fermi Theory as a Matched Low-Energy Layer}

\begin{proposition}[Tree-level matching to Fermi theory]
\label{prop:sm-fermi-tree-matching}
Let \(J^\pm_\mu\) be the charged left-handed weak currents normalized so that
the massive charged-vector part of the electroweak EFT is
\[
  \mathcal L_{\rm cc}
  =
  -\frac{g_2}{\sqrt2}\bigl(J^+_\mu W^{+\mu}+J^-_\mu W^{-\mu}\bigr)
  +m_W^2 W^+_\mu W^{-\mu}
  +\cdots ,
\]
where the dots include the \(W^\pm\) kinetic term and interactions.  In the
low-energy EFT for external momenta \(p^2\ll m_W^2\), eliminating \(W^\pm\) at
leading order gives
\[
  \mathcal L_{\rm Fermi}
  =
  -\frac{g_2^2}{2m_W^2}J^+_\mu J^{-\mu}
  +O(p^2/m_W^4).
\]
Therefore, with the current convention used here,
\[
  \mathcal L_{\rm Fermi}
  =
  -\frac{4G_F}{\sqrt2}J^+_\mu J^{-\mu}
  \qquad\Longleftrightarrow\qquad
  \frac{G_F}{\sqrt2}=\frac{1}{2v^2}.
\]
\end{proposition}

\begin{proof}
At tree level and to leading order in derivatives, the charged vector is an
auxiliary variable with algebraic Lagrangian
\[
  m_W^2 W^+_\mu W^{-\mu}
  -a(J^+_\mu W^{+\mu}+J^-_\mu W^{-\mu}),
  \qquad a:=\frac{g_2}{\sqrt2}.
\]
The equations obtained by varying \(W^+\) and \(W^-\) independently are
\[
  m_W^2W^-_\mu=aJ^-_\mu,\qquad
  m_W^2W^+_\mu=aJ^+_\mu .
\]
Substitution gives
\[
  \left(\frac{a^2}{m_W^2}
  -2\frac{a^2}{m_W^2}\right)J^+_\mu J^{-\mu}
  =
  -\frac{g_2^2}{2m_W^2}J^+_\mu J^{-\mu}.
\]
Including the kinetic operator replaces \(m_W^{-2}\) by the low-momentum
expansion of the massive propagator,
\[
  \frac1{m_W^2-p^2}
  =
  \frac1{m_W^2}+O(p^2/m_W^4),
\]
in the local derivative expansion.  The mass formula from
Proposition~\ref{prop:sm-electroweak-stabilizer-masses} gives
\[
  \frac{g_2^2}{2m_W^2}=\frac{2}{v^2}.
\]
Comparing with
\(-4G_FJ^+_\mu J^{-\mu}/\sqrt2\) proves
\(G_F/\sqrt2=1/(2v^2)\).
\end{proof}

The equality is a tree-level matching statement in the electroweak EFT.  QCD
renormalization of hadronic weak operators, electromagnetic corrections, CKM
factors, and matrix elements of the resulting four-fermion operators are
separate parts of the low-energy matching problem.

\section{Muon \texorpdfstring{\(g-2\)}{g-2} as a Hybrid Prediction}

The anomalous magnetic moment of the muon is a useful organizing example
because it is a single low-energy form-factor coordinate whose Standard Model
evaluation uses all parts of
\(\mathfrak D_{\rm SM}\): perturbative QED, perturbative electroweak matching,
and nonperturbative QCD current correlators.

\begin{definition}[Muon Pauli-coordinate observable]
\label{def:sm-muon-gminus-two-observable}
Let \(\ell=\mu\) be the charged-lepton field in the low-energy EFT and let
\(J_{\rm em}^\mu\) be the electromagnetic current normalized so that the muon
has charge \(-e\), \(e>0\).  Since the muon has weak decay channels, the
magnetic moment used here is the real Pauli-operator coordinate in the
low-energy effective Hamiltonian for a one-muon wave packet, equivalently the
real part of the Pauli form factor obtained after matching the
\(\mu\mu\gamma\) vertex and separating absorptive decay operators.  In the
zero-width projection used for this matching, define the Dirac--Pauli form
factors by
\[
  \langle p',\sigma'|J_{\rm em}^\mu(0)|p,\sigma\rangle
  =
  -\frac{\ii e}{(2\pi)^3}
  \bar u_{\sigma'}(p')
  \left[
    \gamma^\mu F_1^\mu(k^2)
    +
    \frac1{m_\mu}S^{\mu\nu}k_\nu F_2^\mu(k^2)
  \right]
  u_\sigma(p),
  \qquad k:=p-p',
  \label{eq:sm-muon-dirac-pauli-form-factors}
\]
with \(S^{\mu\nu}:=-\ii[\gamma^\mu,\gamma^\nu]/4\) in the mostly-plus
spinor convention of Section~\ref{sec:spinor-conventions}.  The charge
normalization is \(F_1^\mu(0)=1\).  The muon anomalous magnetic moment is
\[
  a_\mu:=\frac{g_\mu-2}{2}:=F_2^\mu(0).
  \label{eq:sm-muon-anomalous-magnetic-moment}
\]
\end{definition}

\begin{proposition}[First QED term]
\label{prop:sm-muon-schwinger-term}
In the QED hard-kernel convention of
Chapter~\ref{ch:qed-renormalization-form-factors},
\[
  a_\mu^{\rm QED,1}=\frac{\alpha}{2\pi},
  \qquad
  \alpha:=\frac{e^2}{4\pi}.
  \label{eq:sm-muon-schwinger-term}
\]
\end{proposition}

\begin{proof}
Chapter~\ref{ch:qed-renormalization-form-factors} defines a form-factor basis
\((F,G)\) in which \(F_2^{\rm DP}(0)=-G(0)\).  The one-loop vertex
calculation gives \(G(0)=-\alpha/(2\pi)\).  Substituting this relation into
Definition~\ref{def:sm-muon-gminus-two-observable} gives
\(a_\mu=F_2^\mu(0)=\alpha/(2\pi)\) at first order in QED.  The lepton mass
enters the loop integral but cancels from the zero-transfer dimensionless
coefficient, as shown explicitly in
Proposition~\ref{prop:qed-one-loop-vertex-schwinger}.
\end{proof}

\begin{definition}[Matched decomposition through a specified order]
\label{def:sm-muon-gminus-two-hybrid-decomposition}
Fix a renormalization scheme, an input-parameter chart, and a maximum order in
the electromagnetic, weak, and QCD matching expansions.  The corresponding
Standard Model value for \(a_\mu\) is written
\[
  a_\mu^{\rm SM}
  =
  a_\mu^{\rm QED}
  +a_\mu^{\rm EW}
  +a_\mu^{\rm HVP}
  +a_\mu^{\rm HLbL}
  +a_\mu^{\rm rem}.
  \label{eq:sm-muon-gminus-two-hybrid-decomposition}
\]
Here \(a_\mu^{\rm QED}\) contains the purely leptonic and photon
contributions in the low-energy QED EFT; \(a_\mu^{\rm EW}\) contains diagrams
or matched Pauli-operator coefficients with at least one \(W,Z,h\), or weak
Goldstone/ghost representative in a gauge-fixed computation;
\(a_\mu^{\rm HVP}\) contains contributions in which the photon propagator is
corrected by the gauge-invariant hadronic electromagnetic current two-point
function; \(a_\mu^{\rm HLbL}\) contains the hadronic electromagnetic
four-current contribution coupled to the muon line by photons; and
\(a_\mu^{\rm rem}\) records the specified higher-order remainder and any
separation-scheme convention.
\end{definition}

\begin{controlledapproximation}[Leading electroweak Pauli coefficient]
\label{ca:sm-muon-leading-electroweak}
In the broken-phase electroweak coupling chart, and at leading order in
\(m_\mu^2/m_W^2\), \(m_\mu^2/m_Z^2\), and \(m_\mu^2/m_h^2\), the one-loop weak
contribution to the muon Pauli coordinate is
\[
  a_\mu^{\rm EW,1}
  =
  \frac{G_Fm_\mu^2}{8\sqrt2\pi^2}
  \left[
    \frac53+\frac13(1-4s_W^2)^2
  \right]
  +
  O\!\left(\frac{G_Fm_\mu^4}{m_W^2}\right),
  \label{eq:sm-muon-leading-electroweak}
\]
where \(s_W^2=g_1^2/(g_1^2+g_2^2)\) in the convention
\(Q=T^3+Y\).
\par
The weak contribution is the coefficient of the dimension-five Pauli operator
\[
  \frac{e}{4m_\mu}a_\mu\,
  \bar\mu[\gamma^\mu,\gamma^\nu]\mu\,F_{\mu\nu}
  \label{eq:sm-muon-pauli-operator}
\]
in the low-energy QED EFT obtained after integrating out \(W,Z,h\).  At one
loop the coefficient is obtained by matching the on-shell
\(\mu\mu\gamma\) vertex in the electroweak theory to the same vertex in the
low-energy EFT, expanded in external momenta over \(m_W,m_Z,m_h\).
Gauge-parameter dependence cancels in the sum of vector, Goldstone, and ghost
representatives because the Pauli operator is electromagnetic-gauge invariant
and the matching is performed between physical on-shell matrix elements.

The charged-current \(W\)-sector diagrams give the heavy-vector coefficient
\[
  \frac{G_Fm_\mu^2}{8\sqrt2\pi^2}\,\frac53 .
  \label{eq:sm-muon-w-coefficient}
\]
The neutral-current \(Z\)-sector contribution depends on the vector coupling
of the \(Z\) to the muon.  With
\[
  J_Z^\mu
  =
  \bar\mu\gamma^\mu
  \bigl(g_LP_L+g_RP_R\bigr)\mu,
  \qquad
  g_L=-\frac12+s_W^2,\qquad g_R=s_W^2,
\]
the vector-minus-axial combination entering the leading Pauli coefficient
reduces, after using \(G_F/\sqrt2=1/(2v^2)\), to
\[
  \frac{G_Fm_\mu^2}{8\sqrt2\pi^2}\,
  \frac13(1-4s_W^2)^2 .
  \label{eq:sm-muon-z-coefficient}
\]
The Higgs contribution is suppressed by an additional factor
\(m_\mu^2/m_h^2\) because both the Higgs coupling and the chirality flip are
proportional to the muon Yukawa coordinate.  Adding the charged and neutral
leading terms gives \eqref{eq:sm-muon-leading-electroweak}.  This is a
short-distance matching statement in the electroweak EFT; higher weak loops
and mixed electroweak-hadronic terms are separate entries in
\(a_\mu^{\rm rem}\) unless explicitly included in the specified order.
A complete proof in this monograph requires the explicit one-loop Pauli
projection for the charged-vector, neutral-vector, Goldstone, and ghost
representatives.  Until that appendix-level calculation is supplied, the
displayed formula is used only as the convention-fixed leading matching
coefficient, not as a theorem established here.
\end{controlledapproximation}

\begin{definition}[Hadronic vacuum polarization from QCD current data]
\label{def:sm-muon-hvp-current-data}
Let
\[
  J_{\rm had}^\mu
  =
  \sum_{q} Q_q\,\bar q\gamma^\mu q
  \label{eq:sm-hadronic-em-current}
\]
be the quark part of the electromagnetic current, with the sum taken over the
active quark fields in the chosen low-energy QCD description and with the
current renormalized in the specified regulator.  The connected transverse
current two-point function defines a scalar polarization by
\[
  \ii\int \dd^4x\,\ee^{\ii q\cdot x}
  \langle\vac|\operatorname T J_{\rm had}^\mu(x)J_{\rm had}^\nu(0)|\vac\rangle_{\rm conn}
  =
  (q^\mu q^\nu-q^2\eta^{\mu\nu})\Pi_{\rm had}(q^2)
  +\text{local contact terms}.
  \label{eq:sm-hvp-lorentzian-polarization}
\]
The charge-subtracted coordinate is
\[
  \widehat\Pi_{\rm had}(q^2):=\Pi_{\rm had}(q^2)-\Pi_{\rm had}(0),
  \label{eq:sm-hvp-subtracted-pi}
\]
where the subtraction is part of the electromagnetic input normalization.
\end{definition}

\begin{proposition}[Dispersive form of leading HVP]
\label{prop:sm-muon-hvp-dispersive}
Assume the hadronic current two-point function has a positive spectral
representation and no massless hadronic state in the electromagnetic channel.
Let
\[
  R_{\rm had}(s):=
  \frac{\sigma(e^+e^-\to{\rm hadrons};s)}
       {4\pi\alpha^2/(3s)}
  \label{eq:sm-rhad-definition}
\]
be the spectral density in the standard electromagnetic normalization.  QED
vacuum-polarization effects are removed according to the same input scheme.
Then the leading HVP contribution is
\[
  a_\mu^{\rm HVP,LO}
  =
  \frac13\left(\frac{\alpha}{\pi}\right)^2
  \int_{s_0}^{\infty}\frac{\dd s}{s}\,
  K_{\rm HVP}(s)\,R_{\rm had}(s),
  \label{eq:sm-hvp-dispersive}
\]
where \(s_0\) is the hadronic threshold and
\[
  K_{\rm HVP}(s)
  :=
  \int_0^1\dd x\,
  \frac{x^2(1-x)}
       {x^2+(1-x)s/m_\mu^2}.
  \label{eq:sm-hvp-kernel}
\]
\end{proposition}

\begin{proof}
The photon-line insertion in the one-loop Pauli form-factor calculation gives
the Euclidean-momentum representation
\[
  a_\mu^{\rm HVP,LO}
  =
  \frac{\alpha}{\pi}
  \int_0^1\dd x\,(1-x)
  \bigl[-4\pi\alpha\,\widehat\Pi_{\rm had}(-Q_x^2)\bigr],
  \qquad
  Q_x^2:=\frac{x^2}{1-x}m_\mu^2,
  \label{eq:sm-hvp-feynman-parameter}
\]
with the same Pauli projection as in
Proposition~\ref{prop:qed-one-loop-vertex-schwinger}.  The once-subtracted
spectral representation of the hadronic current correlator is
\[
  -4\pi\alpha\,\widehat\Pi_{\rm had}(-Q^2)
  =
  \frac{\alpha}{3\pi}
  \int_{s_0}^{\infty}\dd s\,
  \frac{Q^2}{s(s+Q^2)}R_{\rm had}(s).
  \label{eq:sm-hvp-subtracted-dispersion}
\]
Substitute \(Q^2=Q_x^2\) into
\eqref{eq:sm-hvp-feynman-parameter}, interchange the positive \(x\)- and
\(s\)-integrals, and use
\[
  (1-x)\frac{Q_x^2}{s+Q_x^2}
  =
  \frac{x^2(1-x)}
       {x^2+(1-x)s/m_\mu^2}.
\]
This gives \eqref{eq:sm-hvp-dispersive}.  The formula is therefore not a
perturbative expansion of QCD: all hadronic dynamics is in the spectral
function \(R_{\rm had}\), equivalently in the current-current correlator of
Definition~\ref{def:sm-muon-hvp-current-data}.
\end{proof}

\begin{definition}[Euclidean lattice HVP coordinate]
\label{def:sm-muon-lattice-hvp-coordinate}
In a Euclidean nonperturbative QCD regulator, define the zero-spatial-momentum
current correlator
\[
  G(t)
  :=
  \frac13\sum_{i=1}^3
  \int_{\mathbb R^3}\dd^3\vec x\,
  \langle J_{\rm had}^i(t,\vec x)J_{\rm had}^i(0,\vec0)\rangle_E ,
  \qquad t\ge0,
  \label{eq:sm-hvp-euclidean-gt}
\]
with the conserved or renormalized lattice current specified as part of
\(\mathfrak D_{\rm QCD}^{\rm lat}\).  The subtracted Euclidean polarization is
then equivalently encoded by
\[
  \widehat\Pi_E(Q^2)
  =
  \int_0^\infty \dd t\,G(t)
  \left[
    t^2-\frac{4}{Q^2}\sin^2\!\left(\frac{Qt}{2}\right)
  \right],
  \label{eq:sm-hvp-time-momentum-pi}
\]
which fixes the normalization convention for \(G(t)\) relative to
\(\widehat\Pi_E\); on a finite lattice the corresponding current
renormalization factor is part of the regulator data.  Combining this
identity with the Euclidean form of \eqref{eq:sm-hvp-feynman-parameter} gives
\[
  a_\mu^{\rm HVP,LO}
  =
  \left(\frac{\alpha}{\pi}\right)^2
  \int_0^\infty\dd t\,\widetilde K_\mu(t)\,G(t),
  \label{eq:sm-hvp-time-momentum}
\]
where \(\widetilde K_\mu\) is the positive QED kernel obtained by performing
the \(Q^2\)- or Feynman-parameter integral after inserting
\eqref{eq:sm-hvp-time-momentum-pi}.  The lattice prediction additionally
requires continuum, infinite-volume, physical-mass, isospin-breaking, and QED
correction data; these are entries of \(\mathcal M\), not optional refinements.
\end{definition}

\begin{definition}[Hadronic light-by-light coordinate]
\label{def:sm-muon-hlbl-coordinate}
The hadronic light-by-light contribution is defined from the connected
Euclidean four-current correlator
\[
  \Pi_{\mu\nu\rho\sigma}(x,y,z)
  :=
  \langle
    J_{\rm had}^\mu(x)J_{\rm had}^\nu(y)
    J_{\rm had}^\rho(z)J_{\rm had}^\sigma(0)
  \rangle_{E,\rm conn}
  \label{eq:sm-hlbl-four-current}
\]
by convolution with the exactly known QED muon-line kernel
\(\mathcal K_{\mu\nu\rho\sigma}(x,y,z;m_\mu)\):
\[
  a_\mu^{\rm HLbL}
  =
  \int \dd^4x\,\dd^4y\,\dd^4z\,
  \mathcal K_{\mu\nu\rho\sigma}(x,y,z;m_\mu)\,
  \Pi^{\mu\nu\rho\sigma}(x,y,z),
  \label{eq:sm-hlbl-master-coordinate}
\]
with the current normalization, contact-term subtraction, finite-volume
treatment, and QED infrared regulator specified in \(\mathcal M\).  Dispersive
descriptions of the same coordinate decompose the four-current spectral data
into intermediate-channel amplitudes; lattice descriptions compute the
four-current Schwinger function directly.  The equality of these descriptions,
when both are available, is a nontrivial matching statement about the same
QCD-defined current correlator.
\end{definition}

\begin{remark}[Data needed for the \(a_\mu\) prediction]
\label{rem:sm-muon-gminus-two-prediction-inputs}
A value assigned to \(a_\mu^{\rm SM}\) is a Standard Model prediction in the
sense of Definition~\ref{def:sm-hybrid-prediction} only after the following
data have been specified:
\begin{enumerate}
\item the lepton-sector QED perturbative order and charge scheme;
\item the electroweak input scheme and the order of weak matching;
\item the HVP source, either a dispersive spectral function with its
      radiative-correction convention or a lattice current correlator with its
      continuum and infinite-volume extrapolations;
\item the HLbL four-current treatment and its contact-term convention;
\item the combination rule for correlations among input parameters and
      systematic uncertainties.
\end{enumerate}
Definition~\ref{def:sm-muon-gminus-two-observable} identifies \(a_\mu\) as a
single Pauli-form-factor coordinate.  The decomposition
\eqref{eq:sm-muon-gminus-two-hybrid-decomposition} is not a decomposition of
the observable itself; it is a decomposition of its computation in a specified
low-energy matching scheme.  The QED and electroweak entries are perturbative
Pauli-operator
matching coefficients.  The HVP and HLbL entries are functionals of
gauge-invariant QCD current correlators, defined either by a spectral
representation or by a finite regulator such as the lattice.  Changing any
item in the list changes the coordinate map from the hybrid Standard Model
definition to the number called \(a_\mu^{\rm SM}\).  Thus the listed data are
necessary for the five requirements in
Definition~\ref{def:sm-hybrid-prediction}.
\end{remark}

\begin{openproblem}[Constructive Standard Model]
\label{op:constructive-standard-model}
Construct a continuum local QFT realizing the Standard Model representation
content, gauge symmetry, anomaly cancellation, BEH phase, QCD confinement
data, and weak interactions with positivity, locality, covariance, and a
controlled relation to the observed parameter regime.  A solution may be a
single nonperturbative construction or a theorem explaining why the hybrid
definition above has a universal continuum domain for the specified
observables.
No such theorem is assumed in this monograph.
\end{openproblem}
