\chapter{Classical Yang-Mills Theory and Matter Fields}
\label{ch:vol2-classical-yang-mills-matter}
\ConventionsLorentzian


\section{Local Gauge Fields}

This chapter formulates the local classical field theory of nonabelian gauge
fields.  We work first on a spacetime region over which the relevant principal
\(G\)-bundle has been trivialized.  In such a local trivialization, a gauge
field is represented by a Hermitian matrix-valued one-form
\[
  A=A_\mu(x)\,\dd x^\mu,
  \qquad
  A_\mu(x)=A^a_\mu(x)t^a .
\]
The index \(\mu=0,\ldots,D-1\) is a spacetime index.  The index \(a\) labels a
basis of the gauge algebra in the Hermitian-generator convention.  If
\(\mathfrak g_{\mathrm{ah}}=\operatorname{Lie}(G)\) is realized by
anti-Hermitian matrices in a unitary representation, then the Hermitian
generators used here span \(i\,\mathfrak g_{\mathrm{ah}}\).  The associated
anti-Hermitian connection one-form, denoted \(\mathsf A_{\mathrm{ah}}\), is
\(\mathsf A_{\mathrm{ah}}=-\ii A\).  We write
\(\mathfrak g:=i\,\mathfrak g_{\mathrm{ah}}\) for this Hermitian coordinate
space.  Throughout this chapter \([X,Y]\) denotes the ordinary matrix
commutator of Hermitian representatives, while the real compact Lie bracket in
Hermitian coordinates is
\[
  [X,Y]_{\mathrm H}:=-\ii[X,Y].
\]
With this convention,
\begin{equation}
  [t^a,t^b]=\ii f^c{}_{ab}t^c .
  \label{eq:ym-lie-algebra-conventions}
\end{equation}
The real constants \(f^c{}_{ab}\) are the structure constants of
\([\ ,\ ]_{\mathrm H}\) in this basis, with the output Lie-algebra label written
first.  Thus \(f^c{}_{ab}=-f^c{}_{ba}\) and the Jacobi identity is
\[
  f^e{}_{ab}f^d{}_{ec}
  +f^e{}_{bc}f^d{}_{ea}
  +f^e{}_{ca}f^d{}_{eb}=0 .
\]
Formulas below include explicit factors of \(\ii\) so that \(A_\mu\),
\(\zeta\), and \(F_{\mu\nu}\) remain Hermitian matrix fields.

The infinitesimal local transformation with parameter
\(\zeta(x)=\zeta^a(x)t^a\) is
\begin{equation}
  \delta_\zeta A_\mu
  =
  \partial_\mu\zeta-\ii[A_\mu,\zeta].
  \label{eq:ym-infinitesimal-gauge-transformation}
\end{equation}
In components,
\begin{equation}
  \delta_\zeta A^a_\mu
  =
  \partial_\mu\zeta^a+f^a{}_{bc}A^b_\mu\zeta^c,
  \label{eq:ym-infinitesimal-component-transformation}
\end{equation}
The derivative term records the freedom to change local trivialization from
point to point, while the commutator term records the adjoint action of the Lie
algebra on itself.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,2.35) {gauge potential};
    \node at (0,1.20) {\(\displaystyle A_\mu(x)=A^a_\mu(x)t^a\)};
    \draw[blue!65!black,->,thick] (-0.10,1.48)--(-0.82,1.78);
    \node[blue!65!black,align=center] at (-1.25,1.88)
      {spacetime\\index \(\mu\)};
    \draw[purple!75!black,->,thick] (0.18,1.50)--(0.84,1.78);
    \node[purple!75!black,align=center] at (1.25,1.88)
      {Lie-algebra\\index \(a\)};
    \draw[orange!80!black,->,thick] (1.10,1.08)--(1.82,0.78);
    \node[orange!80!black,align=center] at (2.16,0.55)
      {basis generator};

    \begin{scope}[shift={(4.4,0)}]
      \node[font=\small] at (1.65,2.35) {infinitesimal covariance};
      \draw[rounded corners=4pt,thick] (0,0.40) rectangle (3.30,1.84);
      \node at (1.65,1.40)
        {\(\displaystyle
        \delta_\zeta A_\mu=\partial_\mu\zeta-\ii[A_\mu,\zeta]\)};
      \node[align=center,blue!65!black] at (0.93,0.76)
        {change of\\local frame};
      \node[align=center,purple!75!black] at (2.45,0.76)
        {adjoint\\rotation};
      \draw[blue!65!black,decorate,decoration={brace,mirror,amplitude=4pt}]
        (0.56,1.18)--(1.40,1.18);
      \draw[purple!75!black,decorate,decoration={brace,mirror,amplitude=4pt}]
        (1.70,1.18)--(2.74,1.18);
    \end{scope}

    \begin{scope}[shift={(9.45,0)}]
      \node[font=\small] at (1.45,2.35) {Lie closure};
      \draw[rounded corners=4pt,thick] (0.0,0.50) rectangle (2.90,1.75);
      \node at (1.45,1.33)
        {\(\displaystyle [t^a,t^b]=\ii f^c{}_{ab}t^c\)};
      \node[align=center] at (1.45,0.88)
        {the constants \(f^c{}_{ab}\)\\encode the bracket};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The local gauge field carries a spacetime vector index and a
  Lie-algebra index.  The infinitesimal transformation combines the derivative
  of the local parameter with the adjoint action of the Lie algebra.}
  \label{fig:ym-index-and-infinitesimal-data}
\end{figure}

The finite transformation is most cleanly stated using a smooth
\(G\)-valued function \(g(x)\).  With
\[
  g(x)=\exp\{\ii\zeta^a(x)t^a\},
\]
define
\begin{equation}
  A_\mu'
  =
  gA_\mu g^{-1}+\ii g\,\partial_\mu g^{-1}.
  \label{eq:ym-finite-gauge-transformation}
\end{equation}
The Baker-Campbell-Hausdorff formula
\[
  \log(\ee^X\ee^Y)
  =
  X+Y+\frac12[X,Y]
  +\frac1{12}[X,[X,Y]]
  -\frac1{12}[Y,[X,Y]]+\cdots
\]
shows how products of exponentials are controlled by the Lie bracket.  The
identity
\[
  \ee^X Y\ee^{-X}
  =
  Y+[X,Y]+\frac1{2!}[X,[X,Y]]+\cdots
\]
then shows that conjugation by \(g(x)\) maps \(\mathfrak g\) to
\(\mathfrak g\).  Thus \eqref{eq:ym-finite-gauge-transformation} produces
another local representative \(A_\mu'=A_\mu^{\prime a}t^a\).

The same formula is also the transition law between local trivializations.
If \(U_i\) and \(U_j\) are two coordinate patches and
\(h_{ji}:U_i\cap U_j\to G\) is the transition function, then their local
connection representatives satisfy
\[
  A_\mu^{(j)}
  =
  h_{ji}A_\mu^{(i)}h_{ji}^{-1}
  +\ii h_{ji}\partial_\mu h_{ji}^{-1}.
\]
Thus the local field \(A_\mu\) is not itself a globally defined tensor on a
nontrivial bundle.  The curvature constructed below does glue covariantly:
\[
  F_{\mu\nu}^{(j)}=h_{ji}F_{\mu\nu}^{(i)}h_{ji}^{-1}.
\]
The local formulas therefore describe a connection on a principal bundle; they
do not assume that the bundle is globally trivial.

\section{Curvature and the Yang-Mills Action}

Define the covariant differential operator
\begin{equation}
  \nabla_\mu:=\partial_\mu-\ii A_\mu .
  \label{eq:ym-adjoint-covariant-operator}
\end{equation}
The gauge coupling has been absorbed into the connection in this chapter:
if \(g_{\rm el}\) denotes the Abelian charge called \(g\) in
Chapter~\ref{ch:qed-external-states}, then relative to the QED convention
\(\partial_\mu-\ii g_{\rm el} A_\mu^{\rm QED}\) the present connection is
\(A_\mu^{\rm here}=g_{\rm el}A_\mu^{\rm QED}\).  With this choice the
curvature and the action carry the coupling in the Yang--Mills kinetic term
rather than in the covariant derivative.  In the \(U(1)\) specialization,
\begin{equation}
  A_\mu^{\rm here}=g_{\rm el}A_\mu^{\rm QED},
  \qquad
  F_{\mu\nu}^{\rm here}
  =
  \partial_\mu A_\nu^{\rm here}-\partial_\nu A_\mu^{\rm here}
  =
  g_{\rm el}F_{\mu\nu}^{\rm QED}.
  \label{eq:ym-qed-curvature-rescaling}
\end{equation}
This is a change of field normalization, not a new physical assumption; every
occurrence of \(F_{\mu\nu}\) below is in the coupling-absorbed Yang--Mills
normalization.
The finite transformation law \eqref{eq:ym-finite-gauge-transformation} is
equivalent to
\begin{equation}
  \nabla_\mu'=g\nabla_\mu g^{-1},
  \label{eq:ym-covariant-operator-transformation}
\end{equation}
where the factors are multiplied as differential operators.  The curvature is
\begin{equation}
  F_{\mu\nu}
  :=
  \ii[\nabla_\mu,\nabla_\nu]
  =
  \partial_\mu A_\nu-\partial_\nu A_\mu-\ii[A_\mu,A_\nu]
  =
  F^a_{\mu\nu}t^a .
  \label{eq:ym-field-strength-definition}
\end{equation}
It follows from \eqref{eq:ym-covariant-operator-transformation} that
\begin{equation}
  F'_{\mu\nu}=gF_{\mu\nu}g^{-1}.
  \label{eq:ym-curvature-covariance}
\end{equation}

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,2.55) {connection};
    \draw[rounded corners=4pt,thick] (-1.5,0.65) rectangle (1.5,1.75);
    \node at (0,1.20) {\(\nabla_\mu=\partial_\mu-\ii A_\mu\)};

    \draw[-{Latex[length=2mm]},thick] (1.75,1.20)--(3.05,1.20);
    \node[align=center] at (2.40,1.60) {commutator};

    \node[font=\small] at (4.55,2.55) {curvature};
    \draw[rounded corners=4pt,thick] (3.10,0.65) rectangle (5.95,1.75);
    \node at (4.53,1.20) {\(F_{\mu\nu}=\ii[\nabla_\mu,\nabla_\nu]\)};

    \draw[-{Latex[length=2mm]},thick] (6.18,1.20)--(7.45,1.20);
    \node[align=center] at (6.82,1.60) {invariant\\trace};

    \node[font=\small] at (9.05,2.55) {action density};
    \draw[rounded corners=4pt,thick] (7.55,0.65) rectangle (10.55,1.75);
    \node at (9.05,1.20)
      {\(\displaystyle -\frac1{4g_{\mathrm{YM}}^2}\operatorname{tr}(F_{\mu\nu}F^{\mu\nu})\)};

    \draw[purple!75!black,thick,-{Latex[length=2mm]}]
      (0,0.45) .. controls (1.6,-0.55) and (3.2,-0.55) .. (4.55,0.45);
    \node[purple!75!black,align=center] at (2.35,-0.78)
      {\(\nabla_\mu'=g\nabla_\mu g^{-1}\) implies
      \(F_{\mu\nu}'=gF_{\mu\nu}g^{-1}\)};
    \node[blue!65!black,align=center] at (9.05,-0.50)
      {cyclicity removes the conjugation};
  \end{tikzpicture}
  }
  \caption{The Yang-Mills field strength is the curvature of the covariant
  differential operator.  Its conjugation law is converted into a gauge-invariant
  action density by an invariant trace.}
  \label{fig:ym-curvature-action-chain}
\end{figure}

An invariant trace is a symmetric bilinear form on \(\mathfrak g\), written in
a chosen representation as
\begin{equation}
  \operatorname{tr}(t^a t^b)=\kappa^{ab}.
  \label{eq:ym-trace-kappa}
\end{equation}
Cyclicity gives
\[
  \operatorname{tr}([t^a,t^b]t^c)
  =
  \operatorname{tr}(t^a[t^b,t^c])
\]
and, equivalently,
\[
  \operatorname{tr}(gXg^{-1}gYg^{-1})
  =
  \operatorname{tr}(XY)
\]
for Lie-algebra-valued \(X,Y\).  It also gives the invariant tensor
\begin{equation}
  f_{abc}:=\kappa^{ad}f^d{}_{bc}
  =
  -\ii\,\operatorname{tr}\bigl(t^a[t^b,t^c]\bigr),
  \label{eq:ym-fabc-lowered}
\end{equation}
which is completely antisymmetric in \(a,b,c\).  Here
\(\kappa^{ab}:=\operatorname{tr}(t^a t^b)\), while \(\kappa_{ab}\) denotes the
inverse matrix:
\begin{equation}
  \kappa_{ac}\kappa^{cb}=\delta_a{}^b .
  \label{eq:ym-kappa-inverse}
\end{equation}
Consequently the output-first structure constants are recovered from the
trace-lowered constants by
\begin{equation}
  f^a{}_{bc}=\kappa_{ad}f_{dbc}.
  \label{eq:ym-structure-constant-raising}
\end{equation}
The trace normalization used for the Yang--Mills coupling below is
\(\kappa^{ab}=\delta^{ab}\) unless another normalization is explicitly
declared.  In that convention \(f_{abc}=f^a{}_{bc}\) after the obvious
identification of labels.  If instead one uses the common fundamental-matrix
normalization \(\operatorname{tr}(t^a t^b)=\frac12\delta^{ab}\), then
\(\kappa_{ab}=2\delta_{ab}\) and
\[
  f^a{}_{bc}=2\delta_{ad}f_{dbc}.
\]
This factor of \(2\) is a normalization consequence of the chosen trace pairing,
not a separate convention for the Lie bracket.

The Lorentzian Yang-Mills Lagrangian density is
\begin{equation}
  \mathcal L_{\mathrm{YM}}
  =
  -\frac1{4g_{\mathrm{YM}}^2}
  \operatorname{tr}(F_{\mu\nu}F^{\mu\nu}).
  \label{eq:ym-lagrangian-density}
\end{equation}
With the default normalization \(\operatorname{tr}(t^a t^b)=\delta^{ab}\), this
is
\[
  \mathcal L_{\mathrm{YM}}
  =
  -\frac1{4g_{\mathrm{YM}}^2}F^a_{\mu\nu}F^{a\mu\nu}.
\]
Changing the invariant bilinear rescales the component expression and hence the
coordinate called \(g_{\mathrm{YM}}\); the trace-form equation above is the
definition of the coupling convention used in this monograph.
For a unitary Lorentzian quantum theory built from these fields, the invariant
quadratic form on the gauge algebra is chosen positive definite.  Equivalently,
the local gauge algebra is a compact reductive Lie algebra: a direct sum of
abelian \(\mathbb R\) factors and compact simple factors.  The compact simple
factors are the simple Lie algebras of Cartan type
\[
  A,\ B,\ C,\ D,\ E,\ F,\ G .
\]
For these algebras one may choose a linear basis in which
\[
  \kappa^{ab}=\operatorname{tr}(t^a t^b)=\delta^{ab},
\]
which is the normalization used unless another normalization is explicitly
declared.

\begin{remark}[Status of four-dimensional quantum Yang--Mills]
\label{rem:four-dimensional-yang-mills-status}
The construction in this chapter is classical and local.  It supplies the
connection, curvature, invariant action, and matter couplings used later in
perturbation theory and in regulated path integrals.  For a compact simple
gauge group in four Euclidean dimensions, the nonperturbative construction of
pure Yang--Mills theory on \(\mathbb R^4\), together with a proof that the
resulting quantum theory has a positive mass gap, is the Yang--Mills
existence and mass-gap problem of the Clay Millennium list.  A theorem at
that level would construct continuum correlation functions or an equivalent
operator-algebraic theory satisfying the standard structural axioms and would
prove that the spectrum above the vacuum begins at a strictly positive mass.

Consequently, the local Lagrangian \eqref{eq:ym-lagrangian-density}, its
BRST gauge-fixed perturbation theory, and finite-cutoff lattice
regularizations are different mathematical objects from a completed
four-dimensional continuum quantum Yang--Mills theory with mass gap.  Whenever
later chapters use a mass gap, glueball spectrum, or confinement property in
four-dimensional Yang--Mills, that input is explicitly a nonperturbative
hypothesis, a lattice-supported expectation, or a statement inside a regulated
approximation until the continuum existence problem has been solved.
The finite-cutoff lattice formulation itself is developed in
Chapter~\ref{ch:lattice-yang-mills-nonperturbative-definition}, where the
continuum-limit hypothesis and its relation to covariant perturbation theory
are stated separately.
\end{remark}

The basic example \(SU(2)\) can be written in the default trace convention as
\[
  t^a=\frac{\sigma^a}{\sqrt2},\qquad a=1,2,3,
\]
where \(\sigma^a\) are the Pauli matrices.  Then
\(\operatorname{tr}(t^at^b)=\delta^{ab}\) and
\[
  [t^a,t^b]=\ii\sqrt2\,\epsilon^{c}{}_{ab}t^c,
  \qquad
  f^c{}_{ab}=\sqrt2\,\epsilon^{c}{}_{ab}.
\]
The frequently used basis \(\sigma^a/2\) is the alternate half-trace
coordinate system discussed above; it is not the default normalization of this
monograph.  More generally,
the Hermitian coordinate space \(i\,\mathfrak{su}(N)\) is represented by
traceless Hermitian \(N\times N\) matrices, and
\[
  SU(N)=\{g\in U(N):\det g=1\}.
\]

In four spacetime dimensions there is another local gauge-invariant Lorentz
scalar of power-counting dimension four, written in the trace convention just
fixed.\footnote{With \(F_{\mu\nu}=F^a_{\mu\nu}t^a\) and the default
\(\operatorname{tr}(t^at^b)=\delta^{ab}\),
\(\operatorname{tr}(F_{\mu\nu}F_{\rho\sigma})
=F^a_{\mu\nu}F^a_{\rho\sigma}\).  If one instead uses generators normalized by
\(\operatorname{tr}(t^at^b)=\frac12\delta^{ab}\), the components of the same
matrix-valued curvature and the component form of the coupling must be rescaled
accordingly.  The trace-form equations are the convention-independent
statements.  Here and below the trace is the actual matrix trace in the chosen
representation applied to the matrix-valued curvature \(F\); it is not an
independent component contraction that may be rescaled while \(F\) is held
fixed.}
\begin{equation}
  \Delta\mathcal L_\theta
  =
  \frac{\theta}{32\pi^2}
  \epsilon^{\mu\nu\rho\sigma}
  \operatorname{tr}(F_{\mu\nu}F_{\rho\sigma}).
  \label{eq:ym-theta-density}
\end{equation}
With \(\operatorname{tr}(t^at^b)=\delta^{ab}\), this equals
\[
  \frac{\theta}{32\pi^2}
  \epsilon^{\mu\nu\rho\sigma}F^a_{\mu\nu}F^a_{\rho\sigma}.
\]
Authors using a basis with
\(\operatorname{tr}(t^at^b)=\frac12\delta^{ab}\) write the same trace-form
density with half this component coefficient, because their components of
\(F_{\mu\nu}\) are rescaled relative to the basis used here.
The normalization is fixed by the Chern--Weil integer.  For an \(SU(N)\)
bundle over an oriented compact four-manifold \(M\), with
\[
  F=\frac12F_{\mu\nu}\,\dd x^\mu\wedge\dd x^\nu,
\]
the number
\begin{equation}
  Q(P,A)
  :=
  \frac1{8\pi^2}\int_M\operatorname{tr}(F\wedge F)
  =
  \frac1{32\pi^2}
  \int_M\dd^4x\,
  \epsilon^{\mu\nu\rho\sigma}
  \operatorname{tr}(F_{\mu\nu}F_{\rho\sigma})
  \in\mathbb Z
  \label{eq:ym-theta-instanton-number}
\end{equation}
is the integral characteristic number determined by the second Chern class,
with the overall sign fixed by the Hermitian curvature convention.  For
finite-action \(SU(N)\) fields on \(\mathbb R^4\), compactification by the
point at infinity gives the same integer; equivalently, it is the degree of
the limiting pure gauge map \(S^3_\infty\to SU(N)\).
This density is odd under parity and time reversal.  In local coordinates it
is a total derivative, so it does not change the local Euler-Lagrange equations
under boundary conditions that keep the boundary variation fixed.  Perturbation
theory around the trivial gauge-field sector consequently has no local Feynman
vertex associated with \eqref{eq:ym-theta-density}.  Finite-action gauge fields
may approach nontrivial pure gauges at spatial infinity, and then the spacetime
integral of \eqref{eq:ym-theta-density} equals \(\theta Q(P,A)\).  Thus
\(\theta\) enters through global topological sectors of the gauge field, and
the Lorentzian path-integral phase is invariant under
\(\theta\mapsto\theta+2\pi\).  The explicit \(Q=1\) \(SU(2)\) BPST
representative, its action saturation, zero modes, and the resulting
fermionic instanton vertex are constructed in
Section~\ref{sec:bpst-instanton-thooft-vertex}.

\section{Higher-Form Symmetry Data in Gauge Theory}
\label{sec:higher-form-symmetry-data-gauge-theory}

Ordinary internal global symmetries act on local operators.  Gauge theories
also contain symmetries whose charged objects are extended operators.  The
basic example needed later is the center one-form symmetry of pure
Yang--Mills theory.  We record the definition here because it fixes the
meaning of phrases such as center charge, screening, string breaking, and
Wilson-loop order parameter in the QCD chapters.

\begin{definition}[Invertible \(p\)-form symmetry, operational form]
\label{def:invertible-p-form-symmetry}
Let the spacetime dimension be \(D\), and let \(A\) be an abelian group.  An
invertible \(p\)-form global symmetry with group \(A\), in the sector of the
theory under discussion, consists of topological operators
\[
  U_a(\Sigma^{D-p-1}),
  \qquad a\in A,
\]
assigned to closed oriented submanifolds \(\Sigma^{D-p-1}\) of codimension
\(p+1\), together with \(p\)-dimensional charged operators
\(\mathcal W_q(C^p)\).  The operators satisfy
\[
  U_a(\Sigma)U_b(\Sigma)=U_{a+b}(\Sigma),
  \qquad
  U_0(\Sigma)=1,
\]
and are invariant under smooth deformations of \(\Sigma\) that do not cross
charged insertions.  If \(\Sigma\) links \(C\), the symmetry action is
\begin{equation}
  U_a(\Sigma)\,\mathcal W_q(C)\,U_a(\Sigma)^{-1}
  =
  \chi_q(a)^{\operatorname{Lk}(\Sigma,C)}
  \mathcal W_q(C),
  \label{eq:p-form-symmetry-linking-action}
\end{equation}
where \(\chi_q:A\to U(1)\) is the character carried by
\(\mathcal W_q\), and \(\operatorname{Lk}(\Sigma,C)\) is the oriented linking
number when the dimensions are complementary as above.
\end{definition}

For \(p=0\), the topological operator is codimension one and separates space
into two regions; moving it past a local operator implements the ordinary
global symmetry action on that local operator.  For \(p=1\) in four
dimensions, \(U_a(\Sigma)\) is a surface operator, and the charged objects are
line operators.  Invertibility means that the symmetry operators have
inverses.  More general topological defects may have non-group-like fusion;
those categorical and noninvertible structures are part of the later global
structure volume, after line and surface operators have been developed as
QFT observables.

Consider pure \(SU(N)\) Yang--Mills theory with no dynamical matter in
representations of nonzero \(N\)-ality.  The center is
\[
  Z(SU(N))=\{ \ee^{2\pi\ii m/N}\mathbf 1_N:m\in\mathbb Z_N\}.
\]
All local gauge-invariant operators built from \(F_{\mu\nu}\), covariant
derivatives, and adjoint fields are neutral under this center, because the
adjoint representation has zero \(N\)-ality.  A Wilson line in a representation
\(R\),
\[
  W_R(C)=\operatorname{tr}_R P
  \exp\left\{\ii\int_C A_\mu\,\dd x^\mu\right\},
\]
has center charge \(k_R\in\mathbb Z_N\), defined by
\[
  z\,\mathbf 1_N\hbox{ acts on }R\hbox{ as }z^{k_R}.
\]
The electric center one-form symmetry is the \(\mathbb Z_N\) one-form
symmetry whose charged operators are precisely these Wilson lines.

One concrete construction of the symmetry surface is as follows.  Let
\(\Sigma\) be a closed oriented two-surface in four-dimensional spacetime, and
choose a three-chain \(V\) with \(\partial V=\Sigma\) locally.  Across \(V\),
perform a gauge transformation which is single-valued up to the center element
\(z_m=\ee^{2\pi\ii m/N}\mathbf 1_N\).  Adjoint fields and the connection in
the adjoint bundle are unchanged as local gauge-invariant data; a Wilson line
crossing \(V\) acquires the center phase seen by its representation.  The
resulting topological surface operator \(U_m(\Sigma)\) obeys
\begin{equation}
  U_m(\Sigma)W_R(C)U_m(\Sigma)^{-1}
  =
  \exp\left\{
    \frac{2\pi\ii\,m k_R}{N}\operatorname{Lk}(\Sigma,C)
  \right\}
  W_R(C).
  \label{eq:center-one-form-action-wilson-loop}
\end{equation}
This equation is the definition of the center charge used later.  The
topological character of \(U_m(\Sigma)\) says that deforming \(\Sigma\) while
keeping it away from \(C\) leaves the correlation function unchanged; crossing
\(C\) changes the linking number and produces exactly the displayed phase.

Dynamical matter determines the unbroken subgroup.  Suppose the theory
contains matter fields in representations \(R_\alpha\) of \(SU(N)\), with
\(N\)-alities \(k_\alpha\).  A Wilson line in \(R_\alpha\) can end on a
dynamical local field only when that field is present in the operator algebra.
Then a center surface linked with the line endpoint fails to define a
topological operation on the full set of insertions unless its phase on every
dynamical endpoint is trivial.  The surviving electric one-form symmetry is
\begin{equation}
  A_{\mathrm{el}}
  =
  \{m\in\mathbb Z_N:
    \exp(2\pi\ii m k_\alpha/N)=1
    \hbox{ for every dynamical }R_\alpha\}.
  \label{eq:center-one-form-unbroken-subgroup}
\end{equation}
Thus pure Yang--Mills has \(A_{\mathrm{el}}=\mathbb Z_N\), adjoint matter
preserves the same electric center one-form symmetry, and dynamical
fundamental quarks break it completely.

The behavior of large Wilson loops is a dynamical question, but the symmetry
data determine what can serve as a sharp diagnostic.  In a massive pure
Yang--Mills theory, a Wilson loop with nonzero \(N\)-ality is charged under
\(A_{\mathrm{el}}\), so its area or perimeter behavior distinguishes phases of
the one-form symmetry.  With dynamical fundamental quarks the same loop can be
screened by pair creation; the QCD chapter therefore treats the Wilson loop as
a static-source observable, while the one-form-symmetry order-parameter
interpretation applies to the unbroken center subgroup.

\begin{remark}[Higher groups, noninvertible symmetries, and placement]
\label{rem:categorical-symmetry-forward-pointer}
The definition above covers the invertible higher-form symmetry needed for the
center symmetry of pure Yang--Mills.  More refined global structures include
mixed ordinary/higher-form symmetries organized by higher groups, background
\(p+1\)-form gauge fields and their Postnikov constraints, and noninvertible
topological defects with categorical fusion.  Those structures are QFT data
about extended operators and background couplings, and they are conceptually
distinct from the local gauge redundancy used to coordinatize the field space.
They belong to the later volume on global structure, phases, and extended
operators.  The present gauge-theory volume uses only the definitions required
for Wilson lines, screening, confinement diagnostics, and anomaly statements.
\end{remark}

\section{Matter Representations}

Let \(R\) be a finite-dimensional representation of \(G\) on a vector space
\(V_R\).  The matter fields below use complex \(V_R\); real representations,
such as the adjoint representation, are included by complexification when they
are compared with complex representations.  The representation is a
homomorphism
\begin{equation}
  \rho_R:G\longrightarrow GL(V_R),
  \qquad
  \rho_R(gh)=\rho_R(g)\rho_R(h).
  \label{eq:ym-group-representation}
\end{equation}
Its derived Lie-algebra representation is the linear map
\(\widetilde\rho_R:\mathfrak g\to\operatorname{End}(V_R)\) determined by
\begin{equation}
  \rho_R(\ee^{\ii\zeta^a t^a})
  =
  \ee^{\ii\zeta^a t_R^a},
  \qquad
  t_R^a:=\widetilde\rho_R(t^a).
  \label{eq:ym-lie-algebra-representation}
\end{equation}
The matrices \(t_R^a\) obey
\begin{equation}
  [t_R^a,t_R^b]=\ii f^c{}_{ab}t_R^c.
  \label{eq:ym-representation-commutator}
\end{equation}
For compact \(G\), every finite-dimensional complex representation used in a
unitary theory may be equipped with a \(G\)-invariant Hermitian inner product.
In such a basis the matrices \(t_R^a\) are Hermitian.  This is the matter-field
version of the same Hermitian convention used for the gauge potential.

For \(G=SU(N)\), three representations recur throughout gauge theory:
\[
\begin{array}{c|c|c}
\text{representation} & V_R & \text{infinitesimal generators} \\ \hline
\text{fundamental } \square & \mathbb C^N & t_\square^a=t^a \\
\text{anti-fundamental } \overline{\square} & \mathbb C^N &
  t_{\overline{\square}}^a=-(t^a)^* \\
\text{adjoint} & \mathfrak g_{\mathbb C} &
  t_{\mathrm{adj}}^a(t^b)=\ii f^c{}_{ab}t^c
\end{array}
\]
In the fundamental representation,
\(\rho_{\square}(g)=g\) and \(t_{\square}^a=t^a\).  In the
anti-fundamental representation,
\[
  \rho_{\overline{\square}}(g)=g^*
  \qquad\hbox{complex conjugation, without transposition,}
\]
so differentiating
\(\rho_{\overline{\square}}(\ee^{\ii\zeta^a t^a})
=\ee^{\ii\zeta^a t_{\overline{\square}}^a}\) gives
\(t_{\overline{\square}}^a=-(t^a)^*\).
Here \(\mathfrak g_{\mathbb C}\) denotes the complexification of the real
Hermitian coordinate space.  A real adjoint-valued field is obtained by
imposing the corresponding reality condition on its components.  The finite
adjoint representation is conjugation on the algebra:
\[
  v\mapsto gvg^{-1},
  \qquad
  \rho_{\mathrm{adj}}(g)v=gvg^{-1}.
\]
Its infinitesimal form is consistent with the convention
differentiating \(v\mapsto \ee^{\ii\zeta^a t^a}v\ee^{-\ii\zeta^a t^a}\):
\[
  \delta v
  =
  \ii\zeta^a[t^a,v]
  =
  \ii\zeta^a t_{\mathrm{adj}}^a v .
\]

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,3.05) {\(SU(N)\) representation data};

    \draw[rounded corners=4pt,thick] (-5.90,0.35) rectangle (-2.10,2.50);
    \node[font=\bfseries] at (-4.00,2.17) {fundamental \(\square\)};
    \node at (-4.00,1.67) {\(V=\mathbb C^N\)};
    \node at (-4.00,1.25) {\(\rho_\square(g)=g\)};
    \node at (-4.00,0.83) {\(t_\square^a=t^a\)};

    \draw[rounded corners=4pt,thick] (-1.90,0.35) rectangle (1.90,2.50);
    \node[font=\bfseries\small] at (0,2.17)
      {anti-fundamental \(\overline{\square}\)};
    \node at (0,1.67) {\(V=\mathbb C^N\)};
    \node at (0,1.25) {\(\rho_{\overline{\square}}(g)=g^*\)};
    \node at (0,0.83) {\(t_{\overline{\square}}^a=-(t^a)^*\)};

    \draw[rounded corners=4pt,thick] (2.10,0.35) rectangle (5.90,2.50);
    \node[font=\bfseries] at (4.00,2.17) {adjoint};
    \node at (4.00,1.67) {\(V=\mathfrak g\) or \(\mathfrak g_{\mathbb C}\)};
    \node at (4.00,1.25) {\(\rho_{\mathrm{adj}}(g)v=gvg^{-1}\)};
    \node at (4.00,0.83) {\((t_{\mathrm{adj}}^a)^b{}_{c}=\ii f^b{}_{ac}\)};

    \node[font=\small] at (0,-0.10) {QCD quark indices};
    \draw[rounded corners=4pt,thick] (-3.60,-1.95) rectangle (3.60,-0.45);
    \node at (0,-0.92)
      {\(\psi^{\,i}_{I\alpha}(x)\)};
    \node[blue!65!black,align=center] at (-2.22,-1.47)
      {\(i=1,\ldots,N\)\\color in \(\square\)};
    \node[purple!75!black,align=center] at (0,-1.47)
      {\(I=1,\ldots,N_f\)\\flavor};
    \node[orange!80!black,align=center] at (2.22,-1.47)
      {\(\alpha\)\\spinor};
  \end{tikzpicture}
  }
  \caption{The fundamental, anti-fundamental, and adjoint representations of
  \(SU(N)\) used in Yang-Mills theory and QCD.  The anti-fundamental finite
  action is complex conjugation of the matrix entries, not Hermitian
  conjugation.}
  \label{fig:ym-su-n-representation-data}
\end{figure}

If \(\psi(x)\in V_R\), its finite gauge transformation is
\begin{equation}
  \psi'(x)=\rho_R(g(x))\psi(x).
  \label{eq:ym-matter-finite-transformation}
\end{equation}
The covariant derivative acting on \(\psi\) is
\begin{equation}
  D_\mu\psi
  =
  \partial_\mu\psi-\ii A^a_\mu t_R^a\psi.
  \label{eq:ym-matter-covariant-derivative}
\end{equation}
Using \eqref{eq:ym-finite-gauge-transformation} and
\eqref{eq:ym-group-representation}, one obtains
\begin{equation}
  (D_\mu\psi)'=\rho_R(g)D_\mu\psi.
  \label{eq:ym-matter-covariant-derivative-transform}
\end{equation}
This covariance is the algebraic reason that replacing ordinary derivatives
by \(D_\mu\) produces gauge-invariant matter kinetic terms.
For example, if \((\ ,\ )_R\) is the invariant Hermitian pairing on \(V_R\),
then \((D_\mu\psi,D^\mu\psi)_R\) is gauge invariant for scalar matter.  For
Dirac matter, the gauge index contraction in \(\bar\psi\gamma^\mu D_\mu\psi\)
uses the pairing between \(R^*\) and \(R\).  Mass and Yukawa couplings are
gauge invariant only when their flavor and representation tensors define
intertwiners from the tensor product of fields to the trivial representation.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,2.55) {matter field};
    \draw[rounded corners=4pt,thick] (-1.35,0.65) rectangle (1.35,1.75);
    \node at (0,1.22) {\(\psi(x)\in V_R\)};
    \draw[-{Latex[length=2mm]},thick] (1.55,1.22)--(3.15,1.22);
    \node[align=center] at (2.35,1.65)
      {local\\transformation};

    \draw[rounded corners=4pt,thick] (3.30,0.65) rectangle (6.10,1.75);
    \node at (4.70,1.22) {\(\psi'=\rho_R(g)\psi\)};

    \draw[-{Latex[length=2mm]},thick] (6.30,1.22)--(7.88,1.22);
    \node[align=center] at (7.08,1.65)
      {covariant\\derivative};

    \draw[rounded corners=4pt,thick] (8.05,0.65) rectangle (11.20,1.75);
    \node at (9.62,1.22)
      {\(D_\mu=\partial_\mu-\ii A^a_\mu t_R^a\)};

    \node[blue!65!black,align=center] at (4.70,0.10)
      {\(\rho_R(gh)=\rho_R(g)\rho_R(h)\)};
    \node[purple!75!black,align=center] at (9.62,0.10)
      {\((D_\mu\psi)'=\rho_R(g)D_\mu\psi\)};
  \end{tikzpicture}
  }
  \caption{A matter field transforms in a representation \(R\) of the gauge
  group.  The covariant derivative uses the corresponding Lie-algebra
  generators \(t_R^a\), so \(D_\mu\psi\) transforms in the same representation
  as \(\psi\).}
  \label{fig:ym-matter-representation-covariance}
\end{figure}

\section{The Brout--Englert--Higgs Mechanism}
\label{sec:beh-mechanism}

Gauge redundancy is not a physical global symmetry acting faithfully on the
Hilbert space.  A gauge-theory phase with a scalar expectation value is
therefore not defined by a gauge-variant order parameter.  Its perturbative
description is defined by choosing a coordinate chart on field space near a
minimum of the gauge-invariant potential, then fixing the gauge or passing to
gauge-invariant variables.  The Brout--Englert--Higgs mechanism is the local
quadratic statement that scalar fluctuations tangent to the gauge orbit supply
the longitudinal polarizations of gauge bosons whose generators do not
stabilize the chosen vacuum.  The physical formulation is spectral: the
BRST-invariant or gauge-invariant Hilbert space contains massive vector
particles and scalar excitations, while the would-be Goldstone coordinates are
not independent physical massless particles.

\subsection{Abelian Higgs Model}

Use the coupling-absorbed Abelian convention inherited from
\eqref{eq:ym-qed-curvature-rescaling}.  Thus
\[
  D_\mu\phi=\partial_\mu\phi-\ii A_\mu\phi,
  \qquad
\]
\begin{equation}
  \mathcal L
  =
  -\frac1{4e^2}F_{\mu\nu}F^{\mu\nu}
  +(D_\mu\phi)^*D^\mu\phi
  -\lambda\left(|\phi|^2-\frac{v^2}{2}\right)^2 .
  \label{eq:abelian-higgs-lagrangian-beh}
\end{equation}
The finite gauge transformation is
\[
  A_\mu\mapsto A_\mu+\partial_\mu\alpha,
  \qquad
  \phi\mapsto e^{\ii\alpha}\phi .
\]
Choose a local Cartesian coordinate chart near the minimum
\[
  \phi(x)=\frac1{\sqrt2}\bigl(v+h(x)+\ii\pi(x)\bigr).
\]
The scalar kinetic term gives, to quadratic order,
\begin{equation}
  (D_\mu\phi)^*D^\mu\phi
  =
  \frac12\partial_\mu h\,\partial^\mu h
  +
  \frac12(\partial_\mu\pi-vA_\mu)(\partial^\mu\pi-vA^\mu)
  +O(\text{cubic}).
  \label{eq:abelian-higgs-beh-quadratic}
\end{equation}
The potential gives
\[
  \lambda\left(|\phi|^2-\frac{v^2}{2}\right)^2
  =
  \lambda v^2h^2+O(\text{cubic}),
  \qquad
  m_h^2=2\lambda v^2 .
\]
The gauge field \(A_\mu\) is not canonically normalized; the canonical vector
is \(\mathcal A_\mu=A_\mu/e\).  Therefore the vector mass is
\[
  m_A=ev .
\]
Equation \eqref{eq:abelian-higgs-beh-quadratic} displays the mechanism:
the derivative of the orbit coordinate \(\pi\) appears only through the
combination \(\partial_\mu\pi-vA_\mu\) at quadratic order.

On the open set where \(v+h\neq0\), write polar coordinates
\[
  \phi(x)=\frac{v+H(x)}{\sqrt2}\exp\!\left(\frac{\ii\theta(x)}{v}\right).
\]
The gauge-invariant local vector coordinate
\begin{equation}
  B_\mu
  :=
  A_\mu-\frac1v\partial_\mu\theta
  \label{eq:abelian-higgs-gauge-invariant-vector}
\end{equation}
is invariant under \(A_\mu\mapsto A_\mu+\partial_\mu\alpha\),
\(\theta\mapsto\theta+v\alpha\).  In these coordinates
\[
  (D_\mu\phi)^*D^\mu\phi
  =
  \frac12\partial_\mu H\,\partial^\mu H
  +
  \frac12(v+H)^2B_\mu B^\mu .
\]
This is the unitary-gauge description: the orbit coordinate has been removed
from the local chart and the vector has a mass term.  It is useful for
classical mode counting, but its vector propagator has poor ultraviolet
power-counting behavior.  Perturbative renormalizability is most transparent
in \(R_\xi\) gauges.

The \(R_\xi\) gauge-fixing functional in the coupling-absorbed convention is
\begin{equation}
  \mathcal F_\xi
  =
  \partial_\mu A^\mu+\xi e^2v\pi .
  \label{eq:abelian-higgs-rxi-functional}
\end{equation}
Adding
\[
  \mathcal L_{\rm gf}
  =
  -\frac1{2e^2\xi}\mathcal F_\xi^2
\]
cancels the quadratic mixing term \(v(\partial_\mu A^\mu)\pi\) obtained from
\eqref{eq:abelian-higgs-beh-quadratic} after integrating by parts.  The
quadratic Lagrangian becomes
\[
  \mathcal L_2
  =
  -\frac1{4e^2}F_{\mu\nu}F^{\mu\nu}
  -\frac1{2e^2\xi}(\partial_\mu A^\mu)^2
  +\frac12v^2A_\mu A^\mu
  +\frac12(\partial h)^2-\frac12m_h^2h^2
  +\frac12(\partial\pi)^2-\frac12\xi m_A^2\pi^2 .
\]
In terms of the canonical vector \(\mathcal A_\mu=A_\mu/e\), the propagator is
\begin{equation}
  D_{\mu\nu}(p)
  =
  \frac{-\ii}{p^2-m_A^2+\ii0}
  \left[
    \eta_{\mu\nu}
    -
    \frac{(1-\xi)p_\mu p_\nu}
         {p^2-\xi m_A^2+\ii0}
  \right].
  \label{eq:abelian-higgs-rxi-vector-propagator}
\end{equation}
It decays as \(p^{-2}\) at large momentum for fixed \(\xi\).  The field
\(\pi\) and the ghosts have gauge-dependent mass \(\sqrt{\xi}\,m_A\); they are
not physical particles.  The Faddeev--Popov operator follows by varying
\eqref{eq:abelian-higgs-rxi-functional} under the infinitesimal gauge
transformation.  Since
\[
  \delta_\alpha A_\mu=\partial_\mu\alpha,
  \qquad
  \delta_\alpha\pi=(v+h)\alpha+O(\pi\alpha),
\]
one obtains
\[
  \delta_\alpha\mathcal F_\xi
  =
  \left(\Box+\xi e^2v(v+h)\right)\alpha+O(\pi\alpha).
\]
Thus the ghost action contains the quadratic mass \(\xi m_A^2\) and a local
ghost--ghost--\(h\) vertex.  In an Abelian unbroken gauge theory ghosts
decouple; in the \(R_\xi\) chart of the Abelian Higgs model this field
dependence of the gauge condition makes their BRST role visible.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,2.55) {BEH mode accounting in the Abelian model};
    \draw[rounded corners=4pt,thick] (-5.7,0.35) rectangle (-2.3,1.85);
    \node[align=center] at (-4.0,1.45) {scalar chart};
    \node[align=center] at (-4.0,0.92)
      {\(h\): radial\\\(\pi\): gauge-orbit};
    \draw[-{Latex[length=2mm]},thick] (-2.05,1.10)--(-0.35,1.10);
    \node[align=center] at (-1.20,1.52) {quadratic\\mixing};
    \draw[rounded corners=4pt,thick] (-0.1,0.35) rectangle (3.3,1.85);
    \node[align=center] at (1.6,1.45) {massive vector};
    \node[align=center] at (1.6,0.92)
      {\(A_\mu-\partial_\mu\theta/v\)\\three polarizations};
    \draw[-{Latex[length=2mm]},thick] (3.55,1.10)--(5.25,1.10);
    \node[align=center] at (4.40,1.52) {BRST\\quotient};
    \draw[rounded corners=4pt,thick] (5.5,0.35) rectangle (8.9,1.85);
    \node[align=center] at (7.2,1.45) {physical modes};
    \node[align=center] at (7.2,0.92)
      {massive vector\\radial scalar \(h\)};
    \node[align=center,blue!65!black] at (1.6,-0.15)
      {\(R_\xi\) gauges keep \(\pi,c,\bar c\) with gauge-dependent masses;\\
       unitary gauge removes \(\pi\) locally but hides power counting.};
  \end{tikzpicture}
  }
  \caption{The would-be Goldstone coordinate is a coordinate along the gauge
  orbit.  It supplies the longitudinal polarization in a massive vector
  description, while \(R_\xi\) gauges keep it as an unphysical field paired
  with ghosts by BRST.}
  \label{fig:abelian-beh-mode-accounting}
\end{figure}

\subsection{Nonabelian Mass Matrix and the \texorpdfstring{\(SU(2)\to U(1)\)}{SU(2) to U(1)} Pattern}

Let a scalar field \(\Phi\) take values in a unitary representation \(V_R\) of
a compact gauge group \(G\), and let \(v\in V_R\) be a minimum of a
gauge-invariant potential.  The stabilizer subgroup is
\[
  H=\{g\in G:\rho_R(g)v=v\},
  \qquad
  \mathfrak h=\{X\in\mathfrak g:t_R(X)v=0\}.
\]
The tangent space to the gauge orbit at \(v\) is
\[
  T_v(G\cdot v)
  =
  \{\,\ii t_R(X)v:X\in\mathfrak g\,\}
  \simeq \mathfrak g/\mathfrak h .
\]
It is useful to isolate the finite-dimensional linear map
\begin{equation}
  q_v:\mathfrak g\longrightarrow T_vV_R,
  \qquad
  q_v(X)=\ii t_R(X)v .
  \label{eq:beh-orbit-differential}
\end{equation}
The kernel of \(q_v\) is \(\mathfrak h\).  Thus \(q_v\) is the differential
of the gauge-orbit map at the chosen vacuum, and all local mode counting in
the Brout--Englert--Higgs mechanism is the mode counting of this linear map.
Expanding the scalar kinetic term around \(\Phi=v+\eta\) gives the gauge-boson
mass bilinear
\begin{equation}
  (D_\mu v,D^\mu v)_R
  =
  A_\mu^aA^{b\mu}\,
  (t_R^a v,t_R^b v)_R .
  \label{eq:nonabelian-beh-mass-bilinear}
\end{equation}
The Hermitian matrix
\begin{equation}
  K_{ab}:=\operatorname{Re}(t_R^a v,t_R^b v)_R
  \label{eq:nonabelian-beh-K-matrix}
\end{equation}
is positive semidefinite, and its kernel is exactly the unbroken Lie algebra:
\[
  K_{ab}X^b=0
  \quad\Longleftrightarrow\quad
  t_R(X)v=0
  \quad\Longleftrightarrow\quad
  X\in\mathfrak h .
\]
Equivalently, \(K=q_v^\dagger q_v\), where the adjoint is taken with respect
to the invariant inner product on \(V_R\) and the invariant trace form on
\(\mathfrak g\).

\begin{proposition}[Local BEH mode count]
\label{prop:local-beh-mode-count}
At a smooth vacuum \(v\) with compact stabilizer \(H\), the scalar kinetic
term gives nonzero masses to exactly \(\dim G-\dim H\) gauge bosons.  In
four-dimensional Minkowski space these massive vector particles carry
\(3(\dim G-\dim H)\) physical polarizations.  The scalar coordinates removed
from a local unitary-gauge slice are exactly the tangent coordinates to the
orbit \(G/H\).  The remaining scalar coordinates are coordinates normal to the
gauge orbit, subject to the Hessian of the gauge-invariant potential on a
local slice.
\end{proposition}

\begin{proof}
Choose an \(H\)-invariant orthogonal decomposition of scalar fluctuations
near \(v\),
\[
  T_vV_R
  =
  \operatorname{im} q_v\oplus N_v,
\]
where \(N_v\) is the orthogonal complement of the orbit tangent space.  The
existence of such a complement follows by averaging the Hermitian inner
product over the compact group \(H\).  To quadratic order in fields,
\[
  D_\mu(v+\eta)
  =
  \partial_\mu\eta-q_v(A_\mu)+O(A_\mu\eta),
  \qquad A_\mu=A_\mu^a t^a .
\]
If the orbit component of \(\eta\) is written as \(q_v(\chi)\), then the
quadratic expression depends on \(A_\mu\) and \(\chi\) through
\[
  \partial_\mu q_v(\chi)-q_v(A_\mu)
  =
  q_v(\partial_\mu\chi-A_\mu).
\]
This identity is the precise local statement behind the phrase that the
would-be Goldstone coordinate supplies the longitudinal vector polarization:
the scalar orbit coordinate and the gauge potential appear in the same
gauge-orbit differential.  Since \(\operatorname{rank}q_v=\dim\mathfrak g-
\dim\mathfrak h\), exactly that many vector fields have nonzero mass
eigenvalues.  A massive vector particle in four spacetime dimensions has three
physical polarizations after the BRST or gauge-invariant quotient.  The
normal scalar coordinates in \(N_v\) are not gauge-orbit coordinates; their
quadratic masses are the eigenvalues of the potential Hessian restricted to a
local slice transverse to the orbit.
\end{proof}

After canonical normalization of the gauge kinetic term, the vector
mass-squared matrix is \(2g_{\rm YM}^2K\), with the invariant trace form used
to identify gauge-algebra indices.  Hence precisely
\(\dim G-\dim H\) gauge bosons acquire mass.  The same number of scalar
coordinates tangent to \(G/H\) are gauge-orbit coordinates.  A covariant
\(R_\xi\) gauge keeps these coordinates in the field algebra but gives them
gauge-dependent masses; a unitary-gauge chart removes them locally.

The nonabelian \(R_\xi\) construction is the adjoint of the same orbit
differential.  In a local linear chart, let \(q_v^\dagger\eta\in\mathfrak g\)
be the adjoint projection of the scalar fluctuation onto the broken gauge
directions.  The gauge-fixing functional
\begin{equation}
  \mathcal F_\xi
  =
  \partial^\mu A_\mu+\xi g_{\rm YM}^2 q_v^\dagger\eta
  \label{eq:nonabelian-beh-rxi-functional}
\end{equation}
is the nonabelian analogue of
\eqref{eq:abelian-higgs-rxi-functional}.  Its quadratic cross term cancels
the integration-by-parts mixing from
\(\|\partial_\mu\eta-q_v(A_\mu)\|^2\).  In a basis of broken generators
diagonalizing \(K\), \eqref{eq:nonabelian-beh-rxi-functional} becomes a direct
sum of Abelian \(R_\xi\) gauges, with unphysical scalar and ghost masses
\(\sqrt{\xi}\) times the corresponding vector masses.  The unbroken
\(\mathfrak h\) directions have no \(q_v^\dagger\eta\) term and remain
ordinary massless gauge directions at this stage.

For the adjoint \(SU(2)\) model, take an adjoint scalar
\(\Phi=\Phi^a t^a\) and a vacuum \(\Phi_0=v t^3\).  The stabilizer of
\(\Phi_0\) is the \(U(1)\) generated by \(t^3\).  The directions \(t^1,t^2\)
do not stabilize the vacuum, so the corresponding two gauge bosons are
massive and the \(t^3\) gauge boson remains massless.  In electroweak
\(SU(2)\times U(1)\) with a fundamental scalar doublet, the same mass-matrix
logic leaves the diagonal electromagnetic \(U(1)\) unbroken and gives masses
to the orthogonal charged and neutral vector combinations.  The detailed
electroweak parameter conventions belong to the Standard Model chapter; the
structural statement is already contained in
\eqref{eq:nonabelian-beh-mass-bilinear}.

The perturbative renormalizability statement is also local in field space.
An \(R_\xi\) gauge-fixed action for the shifted fields contains only local
operators of power-counting dimension at most four when the original
Yang--Mills--scalar action is renormalizable.  The propagators have the
large-momentum behavior required for power counting, and the BRST identities
restrict counterterms to gauge-invariant deformations together with
BRST-exact gauge-fixing and field-coordinate terms, as in
Chapter~\ref{ch:volume-ii-18}.  Unitary gauge is a useful reduced coordinate
description of the classical spectrum, but it is not the gauge in which the
renormalizability proof is manifest.

\section{Finite-Energy Gauge-Higgs Solitons}
\label{sec:gauge-higgs-classical-solitons}

The finite-energy configuration space of a nonabelian gauge-Higgs theory has
connected components, asymptotic winding data, and nontrivial critical points
of the energy functional.  A classical soliton in this section is such a
finite-energy critical point, together with its boundary sector and its
localized energy density.  Quantizing small fluctuations about such a critical
point gives a semiclassical expansion with different background data from the
translation-invariant vacuum expansion.

We formulate the discussion in the same Hermitian-generator convention and
trace-form coupling convention used above.  Let \(G\) be compact, let
\(\Phi(x)\) be an adjoint scalar field, and let
\[
  D_i\Phi:=\partial_i\Phi-\ii[A_i,\Phi],
  \qquad
  B_i:=\frac12\varepsilon_{ijk}F_{jk}
\]
on a time slice \(\mathbb R^3\).  The index \(i=1,2,3\) is spatial and
\(\varepsilon_{123}=1\).  A simple gauge-Higgs energy functional is
\begin{equation}
  E[A,\Phi]
  =
  \frac1{2g_{\mathrm{YM}}^2}
  \int_{\mathbb R^3}\dd^3x\,
  \operatorname{tr}\bigl(B_iB_i+D_i\Phi D_i\Phi\bigr)
  +
  \frac1{g_{\mathrm{YM}}^2}
  \int_{\mathbb R^3}\dd^3x\,U(\Phi),
  \label{eq:gauge-higgs-static-energy}
\end{equation}
where \(U(\Phi)\ge 0\) is gauge invariant.  For example, for
\(G=SU(2)\) one may take
\[
  U(\Phi)=\frac\lambda4\bigl(\operatorname{tr}\Phi^2-v^2\bigr)^2,
  \qquad
  v>0,\quad \lambda\ge 0.
\]
The scalar \(\Phi\) and the connection \(A_i\) are in the coupling-absorbed
normalization of this chapter.  If one instead writes the covariant derivative
as \(\partial_i-\ii g_{\mathrm{phys}}A^{\mathrm{phys}}_i\), then
\(A_i=g_{\mathrm{phys}}A_i^{\mathrm{phys}}\); the trace-form energy
\eqref{eq:gauge-higgs-static-energy} is the convention used below.

\begin{definition}[Smooth finite-energy sector]
\label{def:smooth-finite-energy-gauge-higgs-sector}
Fix a vacuum orbit
\[
  \mathcal O_v:=\{\Phi\in\mathfrak g:U(\Phi)=0\}.
\]
In the smooth classical sector used in this section, a static configuration
\((A_i,\Phi)\) is finite energy if the integral
\eqref{eq:gauge-higgs-static-energy} is finite and if, after a gauge choice on
the complement of a compact set,
\[
  U(\Phi(x))\to 0,\qquad
  D_i\Phi(x)\to 0,\qquad
  B_i(x)\to 0
  \quad
  \hbox{fast enough as }|x|\to\infty
\]
to make the surface integrals below finite.  The limiting Higgs field defines
a map
\[
  \Phi_\infty:S^2_\infty\longrightarrow \mathcal O_v .
\]
Two configurations are in the same asymptotic sector if their maps
\(\Phi_\infty\) are homotopic through finite-energy boundary data, modulo
gauge transformations which extend over the bulk.
\end{definition}

A more analytic treatment replaces smooth fields by Sobolev completions and
states the boundary conditions as weighted norm conditions.  The topological
classification and the Bogomolny argument below are already visible in the
smooth sector, where all integrations by parts can be justified by decay and
then extended by completion when the appropriate functional-analytic
hypotheses are imposed.

Let \(\Phi_0\in\mathcal O_v\) and let
\[
  H:=\operatorname{Stab}_G(\Phi_0)
  =
  \{h\in G:h\Phi_0h^{-1}=\Phi_0\}.
\]
For an adjoint orbit \(\mathcal O_v\simeq G/H\), the asymptotic Higgs field is
a map \(S^2_\infty\to G/H\).  Hence possible magnetic sectors are measured by
\(\pi_2(G/H)\).  When \(G\) is simply connected, the long exact homotopy
sequence of
\[
  H\longrightarrow G\longrightarrow G/H
\]
gives
\[
  \pi_2(G/H)\simeq \pi_1(H),
\]
because \(\pi_2(G)=0\) for compact connected Lie groups and
\(\pi_1(G)=0\) for simply connected \(G\).  Thus the magnetic charge is the
winding of the transition function of the unbroken group \(H\) seen at
spatial infinity.

\begin{example}[\(SU(2)\to U(1)\)]
\label{ex:su2-to-u1-monopole-sector}
For \(G=SU(2)\) and a nonzero adjoint vacuum value, the stabilizer is
\(H=U(1)\), so
\[
  \mathcal O_v\simeq SU(2)/U(1)\simeq S^2,
  \qquad
  \pi_2(\mathcal O_v)\simeq \mathbb Z .
\]
The integer is the degree of the normalized asymptotic Higgs field
\[
  \widehat\Phi_\infty:S^2_\infty\to S^2,
  \qquad
  \widehat\Phi_\infty(\hat x)=\Phi_\infty(\hat x)/v
\]
when \(\operatorname{tr}\Phi_\infty^2=v^2\).  This integer is invariant under
gauge transformations which extend over the bulk.  In a gauge where
\(\Phi_\infty\) is made constant separately on the
northern and southern hemispheres of \(S^2_\infty\), the two gauges differ on
the equator by a \(U(1)\)-valued transition function of the same winding.
\end{example}

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,2.62) {finite-energy boundary data};
    \draw[thick] (0,0.92) circle (1.05);
    \node at (0,0.92) {\(S^2_\infty\)};
    \draw[blue!65!black,->,thick] (-0.74,1.66).. controls (-1.48,2.20) and (-2.18,1.56)..(-2.36,0.90);
    \draw[purple!75!black,->,thick] (0.74,0.18).. controls (1.45,-0.40) and (2.10,0.23)..(2.36,0.88);
    \node[blue!65!black,align=center] at (-2.72,0.48)
      {north\\trivialization};
    \node[purple!75!black,align=center] at (2.76,0.48)
      {south\\trivialization};

    \draw[-{Latex[length=2mm]},thick] (1.35,1.08)--(3.42,1.08);
    \node[align=center] at (2.40,1.50)
      {\(\Phi_\infty\)};

    \draw[thick] (4.72,0.92) circle (1.05);
    \node at (4.72,0.92) {\(G/H\)};
    \node[align=center] at (4.72,-0.54)
      {vacuum orbit};

    \draw[-{Latex[length=2mm]},thick] (5.98,1.08)--(8.10,1.08);
    \node[align=center] at (7.05,1.50)
      {homotopy class};

    \draw[rounded corners=4pt,thick] (8.20,0.23) rectangle (11.90,1.92);
    \node at (10.05,1.42) {\(\pi_2(G/H)\)};
    \node at (10.05,0.88) {\(\simeq \pi_1(H)\) if \(G\) is simply connected};
    \node at (10.05,0.46) {magnetic sector};

    \node[align=center] at (0,-0.80)
      {\(D_i\Phi\to0,\ B_i\to0,\ U(\Phi)\to0\)};
  \end{tikzpicture}
  }
  \caption{Finite energy turns spatial infinity into a two-sphere mapped to
  the vacuum orbit.  For \(SU(2)\to U(1)\), the degree of this map is the
  magnetic charge of the smooth monopole sector.}
  \label{fig:gauge-higgs-monopole-boundary-map}
\end{figure}

The Bogomolny bound is the simplest place where the topology enters the
energy in addition to labeling boundary data.  Set \(U=0\).  This is the
Prasad--Sommerfield or BPS limit of the adjoint Higgs model.  Completing the
square gives
\begin{align}
  E[A,\Phi]
  &=
  \frac1{2g_{\mathrm{YM}}^2}
  \int_{\mathbb R^3}\dd^3x\,
  \operatorname{tr}\bigl(B_i\mp D_i\Phi\bigr)^2
  \pm
  \frac1{g_{\mathrm{YM}}^2}
  \int_{\mathbb R^3}\dd^3x\,
  \operatorname{tr}(B_iD_i\Phi).
  \label{eq:bogomolny-square}
\end{align}
The Bianchi identity \(D_iB_i=0\) implies
\[
  \operatorname{tr}(B_iD_i\Phi)
  =
  \partial_i\operatorname{tr}(\Phi B_i).
\]
Therefore
\begin{equation}
  E[A,\Phi]
  \ge
  \frac{4\pi}{g_{\mathrm{YM}}^2}
  |\Gamma_{\mathrm m}(A,\Phi)|,
  \qquad
  \Gamma_{\mathrm m}(A,\Phi)
  :=
  \frac1{4\pi}
  \int_{S^2_\infty}\dd S_i\,\operatorname{tr}(\Phi B_i).
  \label{eq:bogomolny-bound-surface}
\end{equation}
The number \(\Gamma_{\mathrm m}\) is the magnetic central charge seen by the
chosen asymptotic Higgs field.  It is the trace pairing of the Higgs vacuum
with the magnetic cocharacter determined by the \(U(1)\) bundle at infinity.
In a primitive \(SU(2)\to U(1)\) sector one may write
\(\Gamma_{\mathrm m}=v_{\mathrm m}n\), \(n\in\mathbb Z\), where
\(v_{\mathrm m}\) is the trace pairing of \(\Phi_\infty\) with the primitive
magnetic generator.  The value of \(v_{\mathrm m}\) changes if one rescales
the matrix generator basis; the trace-form equation
\eqref{eq:bogomolny-bound-surface} does not.
The inequality is saturated precisely when
\begin{equation}
  B_i=\pm D_i\Phi.
  \label{eq:bogomolny-equation}
\end{equation}
These first-order equations imply the second-order Euler--Lagrange equations
in the BPS limit: applying \(D_i\) to \(B_i=\pm D_i\Phi\) gives
\(D_iD_i\Phi=0\), while the variation of \(A_i\) gives
\(D_jF_{ji}=\ii[\Phi,D_i\Phi]\), which follows by differentiating
\(B_i=\pm D_i\Phi\), using the Bianchi identity, and using
\([D_i,D_j]\Phi=-\ii[F_{ij},\Phi]\).

\begin{proposition}[Topological lower bound for adjoint \(SU(2)\) monopoles]
\label{prop:su2-bogomolny-bound}
Consider smooth finite-energy \(SU(2)\) adjoint-Higgs configurations in the
BPS limit \(U=0\).  Fix the asymptotic Higgs value and the primitive magnetic
cocharacter so that \(\Gamma_{\mathrm m}=v_{\mathrm m}n\) on the sector of
degree \(n\in\mathbb Z\).  Then
\[
  E\ge \frac{4\pi v_{\mathrm m}}{g_{\mathrm{YM}}^2}|n|.
\]
Equality holds if and only if the Bogomolny equation
\eqref{eq:bogomolny-equation} holds with the sign chosen to match the sign of
\(n\).
\end{proposition}

\begin{proof}
Equation \eqref{eq:bogomolny-square} is an algebraic identity.  The square
term is nonnegative because the trace form is positive on the compact gauge
algebra.  The remaining term is a surface integral by the Bianchi identity,
as shown above.  For \(SU(2)\to U(1)\), choose two patches on \(S^2_\infty\)
in which \(\Phi_\infty\) is gauge rotated to a fixed Cartan generator.  The
transition function on the equator is a map \(S^1\to U(1)\) of winding
\(n\).  Stokes' theorem on the two patches expresses
\(\int_{S^2_\infty}\dd S_i\,\operatorname{tr}(\Phi B_i)\) as the trace
pairing of the fixed Higgs value with the first Chern flux of this \(U(1)\)
bundle.  This is \(4\pi v_{\mathrm m}n\) by the definition of the primitive
magnetic normalization.  Equality in the square completion holds exactly when
the square vanishes pointwise.
\end{proof}

An explicit unit monopole is obtained from the hedgehog ansatz.  To avoid
embedding normalization factors in every radial equation, let
\[
  T_a:=\frac{t_a}{\sqrt2}=\frac{\sigma_a}{2},
  \qquad
  [T_a,T_b]=\ii\epsilon^c{}_{ab}T_c,
  \qquad
  \operatorname{tr}(T_aT_b)=\frac12\delta_{ab}.
\]
This \(T_a\) is only an auxiliary \(SU(2)\) basis for the following explicit
solution; the action and coupling convention remain the trace-form convention
\eqref{eq:ym-lagrangian-density}.  With
\(\rho:=u r\), \(u>0\), \(r=|x|\), and \(\hat x^a=x^a/r\), write
\begin{equation}
  \Phi(x)=u\,H(\rho)\,\hat x^aT_a,
  \qquad
  A_i(x)=\frac{1-K(\rho)}{r}\,\epsilon_{aij}\hat x^jT_a .
  \label{eq:hedgehog-monopole-ansatz}
\end{equation}
Regularity at the origin and finite energy at infinity require
\[
  K(0)=1,\qquad H(0)=0,\qquad
  K(\infty)=0,\qquad H(\infty)=1 .
\]
Substitution into \(B_i=D_i\Phi\) gives the ordinary differential equations
\begin{equation}
  K'(\rho)=-K(\rho)H(\rho),
  \qquad
  H'(\rho)=\frac{1-K(\rho)^2}{\rho^2},
  \label{eq:bps-hedgehog-ode}
\end{equation}
where the prime denotes \(\dd/\dd\rho\).  The Prasad--Sommerfield solution is
\begin{equation}
  K(\rho)=\frac{\rho}{\sinh\rho},
  \qquad
  H(\rho)=\coth\rho-\frac1{\rho}.
  \label{eq:prasad-sommerfield-profile}
\end{equation}
Indeed,
\[
  \frac{\dd}{\dd\rho}\left(\frac{\rho}{\sinh\rho}\right)
  =
  -\frac{\rho}{\sinh\rho}\left(\coth\rho-\frac1\rho\right),
\]
and
\[
  \frac{\dd}{\dd\rho}\left(\coth\rho-\frac1\rho\right)
  =
  \frac{1-\rho^2/\sinh^2\rho}{\rho^2}.
\]
Thus \eqref{eq:prasad-sommerfield-profile} solves
\eqref{eq:bps-hedgehog-ode}, is smooth at the origin, approaches the vacuum
orbit at infinity, and saturates the bound
\eqref{eq:bogomolny-bound-surface} in the unit magnetic sector.  Translating
the parameter \(u\) to the default trace basis amounts only to replacing it by
the corresponding \(v_{\mathrm m}\) in the surface charge
\(\Gamma_{\mathrm m}\).

For \(\lambda>0\) the square completion is no longer exact, but the same
topological sector persists.  The second-order radial equations for the
profile functions have a smooth solution with a nonzero core size of order
\(u^{-1}\) times a function of \(\lambda\).  Its mass is larger than the BPS
value.  The nonsingular nonabelian core replaces the singularity of a Dirac
monopole: after gauge fixing
\(\Phi_\infty\) to a constant Cartan generator on patches, the asymptotic
field is described by a \(U(1)\) magnetic bundle, while the full \(SU(2)\)
fields remain smooth at the origin.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,2.75) {Bogomolny structure};
    \draw[rounded corners=4pt,thick] (-5.70,0.55) rectangle (-2.05,2.05);
    \node at (-3.88,1.58) {\(\displaystyle
      E=\frac1{2g^2}\int\operatorname{tr}(B_iB_i+D_i\Phi D_i\Phi)\)};
    \node at (-3.88,1.02) {BPS limit \(U=0\)};

    \draw[-{Latex[length=2mm]},thick] (-1.82,1.30)--(-0.28,1.30);
    \node[align=center] at (-1.05,1.72) {complete\\the square};

    \draw[rounded corners=4pt,thick] (-0.05,0.55) rectangle (3.55,2.05);
    \node at (1.75,1.58)
      {\(\displaystyle \frac1{2g^2}\int\operatorname{tr}(B_i-D_i\Phi)^2\)};
    \node at (1.75,1.02)
      {\(\displaystyle +\frac1{g^2}\int_{S^2_\infty}\operatorname{tr}(\Phi B_i)\dd S_i\)};

    \draw[-{Latex[length=2mm]},thick] (3.78,1.30)--(5.22,1.30);
    \node[align=center] at (4.50,1.72) {topological\\flux};

    \draw[rounded corners=4pt,thick] (5.45,0.55) rectangle (8.95,2.05);
    \node at (7.20,1.58)
      {\(\displaystyle E\ge \frac{4\pi v_{\mathrm m}}{g^2}|n|\)};
    \node at (7.20,1.02)
      {\(B_i=\pm D_i\Phi\) at equality};

    \draw[blue!65!black,thick] (-4.90,-0.35)--(-4.20,-0.35)--(-3.50,-0.35);
    \node[blue!65!black] at (-4.20,-0.72) {finite-energy configuration};
    \draw[purple!75!black,thick,->] (6.25,-0.55)--(7.95,-0.55);
    \node[purple!75!black] at (7.10,-0.92) {magnetic charge sector};
  \end{tikzpicture}
  }
  \caption{In the BPS limit the energy is a nonnegative square plus a surface
  term fixed by the magnetic sector.  The displayed square completion derives
  the monopole mass bound from the finite-energy sector.}
  \label{fig:bogomolny-bound-structure}
\end{figure}

The monopole also has collective coordinates.  Translating the solution gives
three zero modes.  A global \(U(1)\) rotation at infinity gives an angular
coordinate conjugate to electric charge; quantizing this coordinate produces
dyonic sectors.  For \(n\) BPS monopoles, the moduli space is a
\(4|n|\)-dimensional hyperk\"ahler manifold after removing the center-of-mass
and overall phase coordinates in the standard way.  The moduli-space statement
is a theorem about the BPS equations and describes the low-velocity classical
dynamics obtained by restricting the kinetic energy to the zero-mode manifold.
The corresponding quantum construction requires a Hilbert space of
fluctuations and collective-coordinate wave functions, possible fermion zero
modes, and a determination of which
semiclassical sectors survive in the continuum quantum theory being studied.

The same finite-energy logic also produces codimension-two vortices.  In the
Abelian Higgs model on \(\mathbb R^2\), with
\[
  D_i\phi=\partial_i\phi-\ii A_i\phi,\qquad
  B=F_{12},
\]
use the energy per unit length
\begin{equation}
  T[A,\phi]
  =
  \int_{\mathbb R^2}\dd^2x\,
  \left[
    \frac1{2e^2}B^2
    +|D_i\phi|^2
    +\frac{e^2}{2}\bigl(|\phi|^2-v^2\bigr)^2
  \right].
  \label{eq:abelian-higgs-vortex-energy}
\end{equation}
Finite energy implies \(|\phi|\to v\) and \(D_i\phi\to0\) at infinity, so
\(\phi/|\phi|:S^1_\infty\to U(1)\) has winding \(n\).  Equivalently,
\[
  \int_{\mathbb R^2}B\,\dd^2x=2\pi n .
\]
At the critical coupling displayed in
\eqref{eq:abelian-higgs-vortex-energy},
\begin{align}
  T
  &=
  \int_{\mathbb R^2}\dd^2x\,
  \left[
    \frac1{2e^2}\bigl(B\mp e^2(v^2-|\phi|^2)\bigr)^2
    + |(D_1\pm\ii D_2)\phi|^2
  \right]
  \pm v^2\int_{\mathbb R^2}B\,\dd^2x .
  \label{eq:vortex-bogomolny-completion}
\end{align}
Hence
\[
  T\ge 2\pi v^2 |n|,
\]
with equality precisely for the Nielsen--Olesen vortex equations
\[
  (D_1\pm\ii D_2)\phi=0,
  \qquad
  B=\pm e^2(v^2-|\phi|^2).
\]
The vortex is therefore the codimension-two analogue of the monopole:
finite energy converts infinity to a sphere of one lower dimension,
topological winding fixes a flux, and a first-order system appears at a
special value of the scalar potential.

Finite-energy critical points also include unstable saddles.  In
four-dimensional \(SU(2)\)-Higgs theory with a fundamental Higgs doublet, the
finite-energy configuration space contains a saddle point called the
sphaleron.  Its distinguishing data are its position between neighboring vacua
of different Chern--Simons number and its single negative fluctuation mode.
With the Hermitian connection convention of this
chapter, the spatial Chern--Simons functional is
\begin{equation}
  N_{\mathrm{CS}}[A]
  =
  \frac1{8\pi^2}
  \int_{\mathbb R^3}
  \operatorname{tr}\left(
    A\wedge \dd A-\frac{2\ii}{3}A\wedge A\wedge A
  \right),
  \label{eq:spatial-chern-simons-functional}
\end{equation}
defined modulo integers under large gauge transformations when the boundary
conditions compactify space to \(S^3\).  The sphaleron sits, in an appropriate
one-parameter minimax construction, at half-integer Chern--Simons number and
has one negative fluctuation mode.  It controls the height of the classical
barrier between adjacent electroweak vacua; real-time transitions at finite
temperature and tunneling transitions at zero temperature require additional
quantum and statistical analysis.  Thus the sphaleron is an unstable static
saddle of the classical energy.

Solitons enter the theory as critical points or sectors of the same classical
field space, selected by finite-energy boundary conditions and by the topology
of gauge and Higgs data at infinity.  Their quantum interpretation requires a
separate construction of states, collective coordinates, fluctuation
determinants, and possible fermion zero modes.  In gauge theories this
construction is phrased in terms of gauge-invariant observables and
superselection sectors; the local gauge potential itself remains a coordinate
on a redundant description.

\section{QCD Notation and Chiral Masses}

For QCD with gauge group \(SU(N)\), a quark field carries three kinds of
indices:
\[
  \psi^{\,i}_{I\alpha}(x),
  \qquad
  i=1,\ldots,N,\quad
  I=1,\ldots,N_f .
\]
Here \(i\) is the color index in the fundamental representation, \(I\) is a
flavor index, and \(\alpha\) is a spinor index.  The conjugate field
\(\bar\psi\) carries the anti-fundamental color representation.

For these quarks, the local four-dimensional terms that are Poincare
invariant, gauge invariant, and power-counting renormalizable have the gauge
part
\[
  -\frac1{4g_{\mathrm{YM}}^2}\operatorname{tr}(F_{\mu\nu}F^{\mu\nu})
  +
  \frac{\theta}{32\pi^2}
  \epsilon^{\mu\nu\rho\sigma}
  \operatorname{tr}(F_{\mu\nu}F_{\rho\sigma})
\]
and the fermion part
\begin{equation}
  \mathcal L_\psi
  =
  -\bar\psi_I\gamma^\mu D_\mu\psi_I
  -
  m_{IJ}\bar\psi_I\psi_J
  -
  \widetilde m_{IJ}\bar\psi_I\gamma_5\psi_J .
  \label{eq:qcd-fermion-mass-lagrangian}
\end{equation}
The matrices \(m\) and \(\widetilde m\) act on flavor and are Hermitian in a
Hermitian Lorentzian action.

For chiral notation, use the convention fixed in
Chapter~\ref{ch:spinor-fields-fermionic-path-integrals},
\(\gamma_5=-\ii\gamma^0\gamma^1\gamma^2\gamma^3\); with this sign the
left-handed subspace is the \(+1\) eigenspace of \(\gamma_5\):
\[
  P_L=\frac{1+\gamma_5}{2},
  \qquad
  P_R=\frac{1-\gamma_5}{2},
  \qquad
  \psi_{IL}=P_L\psi_I,\quad \psi_{IR}=P_R\psi_I.
\]
Then the fermion Lagrangian can be written
\begin{equation}
  \mathcal L_\psi
  =
  -\bar\psi_{IL}\gamma^\mu D_\mu\psi_{IL}
  -\bar\psi_{IR}\gamma^\mu D_\mu\psi_{IR}
  -M_{IJ}\bar\psi_{IL}\psi_{JR}
  -M^*_{IJ}\bar\psi_{JR}\psi_{IL}.
  \label{eq:qcd-chiral-mass-lagrangian}
\end{equation}
The complex mass matrix \(M\) is the coefficient matrix of the left-right
bilinear in \eqref{eq:qcd-chiral-mass-lagrangian}; \(M^*_{IJ}\) denotes the
entrywise complex conjugate of \(M_{IJ}\).  Equivalently, after substituting
\(\psi_{IL}=P_L\psi_I\), \(\psi_{IR}=P_R\psi_I\), and the corresponding
Dirac-adjoint projectors, \(M\) is defined by the identity
\[
  m_{IJ}\bar\psi_I\psi_J
  +\widetilde m_{IJ}\bar\psi_I\gamma_5\psi_J
  =
  M_{IJ}\bar\psi_{IL}\psi_{JR}
  +M^*_{IJ}\bar\psi_{JR}\psi_{IL}.
\]
Thus \(M\) is the same mass data expressed in the chiral basis.

\begin{figure}[H]
  \centering
  \resizebox{0.92\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,2.35) {Dirac-field mass data};
    \draw[rounded corners=4pt,thick] (-1.45,0.65) rectangle (1.45,1.70);
    \node at (0,1.30) {\(m_{IJ}\bar\psi_I\psi_J\)};
    \node at (0,0.95) {\(\widetilde m_{IJ}\bar\psi_I\gamma_5\psi_J\)};

    \draw[-{Latex[length=2mm]},thick] (1.72,1.17)--(3.10,1.17);
    \node[align=center] at (2.42,1.58) {chiral\\projectors};

    \node[font=\small] at (4.80,2.35) {left and right components};
    \draw[rounded corners=4pt,thick] (3.20,0.65) rectangle (6.40,1.70);
    \node at (4.80,1.36) {\(\psi_{IL}=P_L\psi_I\)};
    \node at (4.80,0.95) {\(\psi_{IR}=P_R\psi_I\)};

    \draw[-{Latex[length=2mm]},thick] (6.65,1.17)--(8.05,1.17);
    \node[align=center] at (7.35,1.58) {complex\\matrix};

    \node[font=\small] at (9.95,2.35) {chiral mass form};
    \draw[rounded corners=4pt,thick] (8.15,0.65) rectangle (11.80,1.70);
    \node at (9.98,1.34) {\(-M_{IJ}\bar\psi_{IL}\psi_{JR}\)};
    \node at (9.98,0.94) {\(-M^*_{IJ}\bar\psi_{JR}\psi_{IL}\)};
  \end{tikzpicture}
  }
  \caption{Scalar and pseudoscalar Hermitian mass matrices can be repackaged
  into a complex matrix coupling left- and right-handed quark fields.}
  \label{fig:qcd-chiral-mass-data}
\end{figure}

This notation prepares two later developments.  Gauge fixing and BRST will
quantize the gauge theory while preserving the gauge-covariant local data
defined here.  Anomalous chiral rotations will relate phases of the mass matrix
to the \(\theta\)-angle.
