\chapter{Infrared Divergences and Inclusive QED}
\label{ch:volume-ii-infrared-inclusive-qed}
\ConventionsLorentzian


\section{Soft Radiation Data}

The scattering constructions developed earlier are simplest when all
asymptotic particles are massive.  Four-dimensional QED has a massless stable
photon, and a charged
particle may emit photons of arbitrarily small energy.  This chapter constructs
the scattering quantities adapted to that long-distance structure.
The ordinary LSZ theorem for a charged external electron would require an
isolated massive charged pole.
Massless QED instead has charged threshold data accompanied by arbitrarily
soft electromagnetic radiation.  The fixed hard charged labels used below are
therefore not exact fixed-photon-number Wigner-particle \(S\)-matrix labels.
They are hard data in an infrared-regulated factorization problem; physical
charged scattering is attached either to detector-inclusive probabilities or
to asymptotic states dressed by their soft electromagnetic fields.

Throughout the chapter, spacetime is \(D=4\) Minkowski space with metric
\[
  \eta_{\mu\nu}=\operatorname{diag}(-1,+1,+1,+1).
\]
Consider a hard transition
\[
  \alpha\longrightarrow\beta
\]
between massive charged particles.  The hard external charged lines are
labelled by \(n\).  Their momenta are \(p_n\), their charges are \(g_n\), and
\[
  p_n^2=-m_n^2,\qquad p_n^0>0 .
\]
The sign
\[
  \eta_n=
  \begin{cases}
    +1,& n\in\beta\quad\hbox{outgoing},\\
    -1,& n\in\alpha\quad\hbox{incoming}
  \end{cases}
\]
records whether the line is outgoing or incoming.  Charge conservation for
the hard process is
\[
  \sum_n \eta_n g_n=0 .
\]
The charge symbols used in the QED chapters are related as follows.
\begin{center}
\begin{tabular}{p{0.24\textwidth}|p{0.12\textwidth}|p{0.54\textwidth}}
context & symbol & meaning \\
\hline
Chapter~\ref{ch:qed-external-states}
  & \(g\)
  & signed Abelian charge in \(D_\mu=\partial_\mu-\ii gA_\mu\);
    for an electron \(g=-e\), with \(e>0\). \\
Chapter~\ref{ch:qed-renormalization-form-factors}
  & \(e,\ e^B\)
  & positive magnitude of the electron charge and its bare counterpart; the
    electron covariant derivative is \(\partial_\mu+\ii eA_\mu\), and
    \(e=e^B Z_A^{1/2}\). \\
this chapter
  & \(g_n\)
  & signed physical charge of the \(n\)-th hard external line; an electron has
    \(g_n=-e\), a positron has \(g_n=+e\), and an external scalar of charge
    \(g\) in the scalar-QED example has \(g_n=g\).
\end{tabular}
\end{center}
Thus \(\eta_n\) records whether the charged line is incoming or outgoing,
whereas \(g_n\) records the physical sign of the charge carried by that line.

A soft photon has momentum
\[
  k^\mu=(\omega,\vec k),\qquad \omega=|\vec k|,
\]
and physical helicity \(h=\pm\).  The detector threshold is denoted by
\(E_T\): photons below this energy are included in the same observed hard
event.  An auxiliary infrared cutoff \(\mu\) will regulate intermediate
formulas, and \(M\) denotes a soft upper scale below which the eikonal
approximation is valid:
\[
  0<\mu<E_T<M .
\]

\begin{definition}[Regulated resolution-inclusive transition probability]
Let \(S^{(\mu)}_{\beta;\{k_r,h_r\}|\alpha}\) denote an infrared-regulated hard
kernel with hard charged configuration \(\beta\), incoming hard charged
configuration \(\alpha\), and additional photons \((k_r,h_r)\).  The
superscript \((\mu)\) records the auxiliary long-distance regulator.  The
object \(S^{(\mu)}\) may be obtained from a regulated LSZ computation, but it
is not assumed to have a regulator-independent fixed-photon-number limit.  The
regulated resolution-inclusive probability at threshold \(E_T\) is the sum
\[
  P^{(\mu)}_{\beta|\alpha}(E_T)
  =
  \sum_{N=0}^{\infty}\frac1{N!}
  \sum_{h_1,\ldots,h_N}
  \int_{\mathcal R_{E_T}}
  \prod_{r=1}^{N}
  \frac{\dd^3\vec k_r}{(2\pi)^3\,2\omega_r}\,
  \left|
    S^{(\mu)}_{\beta;\{k_r,h_r\}|\alpha}
  \right|^2 ,
\]
where \(\mathcal R_{E_T}\) is the detector-degenerate soft region.  At leading
soft logarithmic order one may take \(\mu<\omega_r<E_T\) independently for
each unresolved photon.  Refining the detector model changes finite functions
of \(E_T/M\), but not the cancellation of the regulator \(\mu\).
The regulator-free inclusive probability is not the termwise limit of the
summands at fixed \(N\); it is the ordered limit constructed in
Section~\ref{sec:volume-ii-inclusive-probabilities}.
\end{definition}

\section{External-Line Soft Factors}

In scalar QED, attach an outgoing photon of momentum \(k\) and polarization
vector \(e_h^\mu(k)\) to an outgoing external scalar line of momentum \(p\).
The factor multiplying the hard amplitude is
\[
  (-\ii)\,
  \frac{1}{(p+k)^2+m^2-\ii\epsilon}\,
  \ii g(2p+k)^\mu e^*_{h,\mu}(k).
\]
Since \(p^2+m^2=0\) and \(k^2=0\), the leading term as \(k\to0\) is
\[
  g\,\frac{p\cdot e_h^*(k)}{p\cdot k-\ii\epsilon}.
\]
The denominator is linear in the photon energy.  The soft photon therefore
couples to the charge and momentum of the external line, while the internal
structure of the hard subdiagram is unresolved at leading order.

For the hard transition \(\alpha\to\beta\), the one-soft-photon amplitude is
\[
  \mathcal M_{\beta;k,h|\alpha}
  =
  \mathcal M_{\beta|\alpha}
  \left[
    \sum_n
    \frac{g_n\,p_n\cdot e_h^*(k)}
         {\eta_n p_n\cdot k-\ii\epsilon}
  \right]
  +O(k^0).
\]
The same leading factor is obtained in spinor QED.  The numerator of the
nearly on-shell fermion propagator acts on the external spinor and reduces to
the corresponding momentum factor; spin first enters in subleading soft
terms.
Indeed, for an outgoing fermion line the soft attachment contains
\[
  \left[
  \bar u^\sigma(p)(-g\gamma^\mu)
  \frac{-\ii[-\ii(\not p+\not k)+m]}
       {(p+k)^2+m^2-\ii\epsilon}
  \right]e^*_{h,\mu}(k).
\]
The denominator is \(2p\cdot k-\ii\epsilon+O(k^2)\).  In the numerator,
using the on-shell equation for the external spinor and moving
\(\gamma^\mu\) through the Dirac numerator gives
\[
  \bar u^\sigma(p)\gamma^\mu(-\ii\not p+m)
  =
  \bar u^\sigma(p)(\ii\not p+m)\gamma^\mu
  -2\ii p^\mu\bar u^\sigma(p)
  =
  -2\ii p^\mu\bar u^\sigma(p).
\]
Thus the leading soft limit is
\[
  g\,\frac{p\cdot e_h^*(k)}{p\cdot k-\ii\epsilon}\,
  \bar u^\sigma(p),
\]
the same eikonal factor multiplying the original spinor amplitude.

\begin{theorem}[Weinberg leading soft photon theorem]
\label{thm:weinberg-leading-soft-photon}
Fix an infrared regulator and hard external data for which the hard kernel
\(\mathcal M_{\beta|\alpha}\) is defined and smooth under the addition of one
photon of momentum \(k^\mu=\omega(1,\widehat k)\) with \(\omega\to0^+\), away
from nonsoft singular surfaces.  Then the hard kernel with one additional
outgoing physical photon of helicity \(h\) satisfies
\begin{equation}
  \mathcal M_{\beta;k,h|\alpha}
  =
  S_\gamma^{(0)}(k,h)\,
  \mathcal M_{\beta|\alpha}
  +O(\omega^0),
  \qquad
  S_\gamma^{(0)}(k,h)
  =
  \sum_n
  \eta_n g_n\,
  \frac{p_n\cdot e_h^*(k)}{p_n\cdot k}.
  \label{eq:weinberg-leading-soft-photon}
\end{equation}
The sum runs over all hard charged external lines, with \(\eta_n=+1\) for
outgoing and \(\eta_n=-1\) for incoming hard lines.  The \(i\epsilon\)
prescription in the preceding external-line formula specifies the boundary
value of the regulated kernel; the theorem displays the homogeneous leading
soft factor.
\end{theorem}

\begin{proof}
The scalar and spinor calculations above show that attaching the soft photon
to an external hard line produces the pole
\[
  \eta_n g_n\,
  \frac{p_n\cdot e_h^*(k)}{p_n\cdot k}\,
  \mathcal M_{\beta|\alpha}.
\]
The sign \(\eta_n\) is the crossing sign: an incoming charged line is
equivalently an outgoing line with momentum and charge flow reversed in the
all-outgoing Ward identity.  Attachments to internal lines are less singular
in the uniform soft limit at fixed nonexceptional hard kinematics, because no
internal propagator is forced onto a new mass shell by adding a momentum
\(k\) of order \(\omega\).  Summing the external-line poles gives
\eqref{eq:weinberg-leading-soft-photon}.

Gauge invariance fixes the allowed leading pole.  A physical photon
polarization representative is defined modulo
\[
  e^\mu_h(k)\sim e^\mu_h(k)+c\,k^\mu .
\]
Under this replacement, the leading factor shifts by
\[
  \delta S_\gamma^{(0)}
  =
  c^*
  \sum_n\eta_n g_n .
\]
Thus the unphysical longitudinal polarization decouples precisely when the
hard process obeys charge conservation,
\[
  \sum_n\eta_n g_n=0.
\]
This is the Ward identity for the global \(U(1)\) charge written in the
external hard basis.  Conversely, a nonzero coefficient of the longitudinal
soft pole would mean that the regulated hard kernel had not descended to a
  physical photon helicity amplitude.
\end{proof}

Figure~\ref{fig:weinberg-leading-soft-ward-identity} records this leading-soft
Ward identity: attaching a soft photon to the charged hard external lines gives
the eikonal factor, and the longitudinal part cancels precisely by charge
conservation.

\begin{figure}[!ht]
  \centering
  \begin{tikzpicture}[>=Stealth,scale=0.92,every node/.style={font=\small}]
    \node[draw,rounded corners=2pt,minimum width=2.4cm,minimum height=1.15cm,
      align=center,fill=blue!4] (hard) at (0,0) {hard kernel\\
      \(\mathcal M_{\beta|\alpha}\)};
    \draw[line width=0.75pt,-{Latex[length=2mm]}] (-2.13,0.99)--(hard);
    \draw[line width=0.75pt,-{Latex[length=2mm]}] (-2.13,-0.99)--(hard);
    \draw[line width=0.75pt,-{Latex[length=2mm]}] (2.13,0.99)--(hard);
    \draw[line width=0.75pt,-{Latex[length=2mm]}] (2.13,-0.99)--(hard);
    \node[align=center] at (-3.18,1.45) {\(p_1,g_1,\eta_1\)};
    \node[align=center] at (-3.18,-1.45) {\(p_2,g_2,\eta_2\)};
    \node[align=center] at (3.18,1.45) {\(p_3,g_3,\eta_3\)};
    \node[align=center] at (3.18,-1.45) {\(p_4,g_4,\eta_4\)};
    \draw[decorate,decoration={snake,amplitude=0.75mm,segment length=3.4mm},
      line width=0.8pt,-{Latex[length=2mm]},red!70!black]
      (hard) -- (0,2.15)
      node[above,red!70!black] {\(k,e_h\)};
    \node[align=center,anchor=west] at (3.65,0.58)
      {soft factor\\[-0.2em]
      \(S_\gamma^{(0)}
      =\sum_n\eta_ng_n\frac{p_n\cdot e_h^*}{p_n\cdot k}\)};
    \node[align=center,anchor=west] at (3.65,-0.72)
      {gauge shift \(e_h\mapsto e_h+ck\)\\[-0.2em]
      \(\delta S_\gamma^{(0)}=c^*\sum_n\eta_ng_n=0\)};
  \end{tikzpicture}
  \caption{The leading soft photon factor is the sum of external-line poles.
  Replacing the polarization by a pure gauge vector tests the Ward identity;
  the pole descends to the physical photon helicity space exactly when the hard
  charges are conserved.}
  \label{fig:weinberg-leading-soft-ward-identity}
\end{figure}

\begin{theorem}[Weinberg leading soft graviton theorem]
\label{thm:weinberg-leading-soft-graviton}
Consider an infrared-regulated theory with a massless spin-two gauge field
whose representative is normalized by
\[
  g_{\mu\nu}=\eta_{\mu\nu}+\kappa h_{\mu\nu},
  \qquad
  \mathcal L_{\rm int}
  =
  -\frac{\kappa}{2}h_{\mu\nu}T^{\mu\nu}.
\]
For one additional outgoing graviton of momentum \(k\) and physical
polarization tensor \(\varepsilon_{\mu\nu}^{(\lambda)}(k)\), the leading soft
limit is
\begin{equation}
  \mathcal M_{\beta;k,\lambda|\alpha}^{\rm grav}
  =
  S_{\rm grav}^{(0)}(k,\lambda)\,
  \mathcal M_{\beta|\alpha}
  +O(\omega^0),
  \qquad
  S_{\rm grav}^{(0)}
  =
  \frac{\kappa}{2}
  \sum_n
  \eta_n\,
  \frac{p_n^\mu p_n^\nu
  \varepsilon_{\mu\nu}^{(\lambda)*}(k)}{p_n\cdot k}.
  \label{eq:weinberg-leading-soft-graviton}
\end{equation}
\end{theorem}

\begin{proof}
The stress-tensor vertex on an external massive line contains, at leading
order in the soft momentum, the matrix element of the hard momentum carried by
that line.  For a scalar line, the one-graviton vertex contracted with a
transverse traceless polarization gives
\[
  \kappa\,
  p^\mu p^\nu\varepsilon_{\mu\nu}^*(k)
\]
up to terms less singular in \(k\), while the adjacent nearly on-shell
propagator gives \((2\eta_n p_n\cdot k)^{-1}\).  Their product is the
\((\kappa/2)\eta_n p_n^\mu p_n^\nu\varepsilon_{\mu\nu}^*/(p_n\cdot k)\)
term in \eqref{eq:weinberg-leading-soft-graviton}.  Spin-dependent terms in
the stress tensor contain an additional power of the soft momentum or the
Lorentz generator and therefore begin at subleading order.  Summing over the
hard external lines gives \eqref{eq:weinberg-leading-soft-graviton}.

The physical spin-two polarization is defined modulo the linearized
diffeomorphism shift
\[
  \varepsilon_{\mu\nu}\sim
  \varepsilon_{\mu\nu}
  +k_\mu\xi_\nu+k_\nu\xi_\mu ,
  \qquad
  k\cdot\xi=0
\]
on the null momentum \(k\).  The leading soft factor changes by
\[
  \delta S_{\rm grav}^{(0)}
  =
  \kappa
  \xi_\nu^*
  \sum_n\eta_n p_n^\nu .
\]
This vanishes by hard momentum conservation.  If the coefficient of
\(h_{\mu\nu}T^{\mu\nu}\) depended on the species, the same gauge variation
would contain
\[
  \sum_n\eta_n\kappa_n p_n^\nu ,
\]
which does not vanish for arbitrary allowed hard reactions unless all
\(\kappa_n\) are the same constant.  The leading soft spin-two theorem
therefore packages both the external-line pole and the universality of the
gravitational coupling.
\end{proof}

The photon theorem is the input used in the rest of this chapter.  The
graviton theorem is recorded here because it is the same infrared consequence
of a massless gauge particle and gauge-invariant factorization.  A complete
quantum theory containing the spin-two field requires additional ultraviolet
and nonperturbative input not used in the soft argument.

For \(N\) real soft photons the leading factors multiply:
\[
  \mathcal M_{\beta;\{k_r,h_r\}|\alpha}
  =
  \mathcal M_{\beta|\alpha}
  \prod_{r=1}^{N}
  \left[
    \sum_n
    \frac{g_n\,p_n\cdot e_{h_r}^*(k_r)}
         {\eta_n p_n\cdot k_r-\ii\epsilon}
  \right]
  +O(k_r^0).
\]
For two photons emitted from the same line, the two possible orderings of the
emissions turn the nested eikonal denominators into the product displayed
above.  For an outgoing scalar line this is the elementary identity
\[
\begin{aligned}
  &\frac{g\,p\cdot e_{h_2}^*(k_2)}
        {p\cdot k_2-\ii\epsilon}\,
   \frac{g\,p\cdot e_{h_1}^*(k_1)}
        {p\cdot(k_1+k_2)-\ii\epsilon}
  +(1\leftrightarrow2)
\\
  &\qquad =
  \frac{g\,p\cdot e_{h_1}^*(k_1)}
       {p\cdot k_1-\ii\epsilon}\,
  \frac{g\,p\cdot e_{h_2}^*(k_2)}
       {p\cdot k_2-\ii\epsilon}.
\end{aligned}
\]
Corrections from the non-eikonal parts of the propagators and vertices are
less singular in the soft scaling.  The same combinatorics gives the product
for any finite number of soft photons.

\begin{remark}[Subleading soft photons]
The preceding product formula uses only the leading eikonal term.  For a
single additional photon, under the perturbative hypotheses of massive charged
external particles, a smooth hard amplitude near \(k=0\), and a fixed
infrared regulator, the Low soft expansion has the form
\[
  \mathcal M_{\beta;k,h|\alpha}
  =
  \left(S_h^{(0)}(k)+S_h^{(1)}(k)\right)
  \mathcal M_{\beta|\alpha}
  +O(\omega),
\]
where
\[
  S_h^{(0)}(k)
  =
  \sum_n
  \frac{g_n\,p_n\cdot e_h^*(k)}
       {\eta_n p_n\cdot k}
\]
is the leading factor used above, and the subleading operator is
\[
  S_h^{(1)}(k)
  =
  -\ii\sum_n
  \frac{g_n\,e^*_{h,\mu}(k)k_\nu J_n^{\mu\nu}}
       {\eta_n p_n\cdot k}.
\]
Here \(J_n^{\mu\nu}\) is the total Lorentz generator acting on the \(n\)-th
hard external datum.  For a scalar external line its orbital part is
\[
  J_{n,\mathrm{orb}}^{\mu\nu}
  =
  -\ii\left(
    p_n^\mu\frac{\partial}{\partial p_{n,\nu}}
    -
    p_n^\nu\frac{\partial}{\partial p_{n,\mu}}
  \right),
\]
and for spinning particles one adds the spin matrix
\(\Sigma_n^{\mu\nu}\).  The formula is a distributional statement on the
mass shell and momentum-conservation surface: the differential operator acts
on the hard amplitude together with its conservation delta function, or
equivalently on a chosen smooth local representative before restricting to the
physical surface.

The subleading factor is homogeneous of degree \(\omega^0\), whereas
\(S_h^{(0)}\) is homogeneous of degree \(\omega^{-1}\).  Consequently the
real-emission integral built from \(S_h^{(0)}S_h^{(0)*}\) contains
\(\dd\omega/\omega\), while the interference between \(S_h^{(0)}\) and
\(S_h^{(1)}\) contains \(\dd\omega\) and is finite at the lower endpoint in
the massive case.  This is why the Bloch--Nordsieck logarithm and its Poisson
exponentiation are governed by \(S^{(0)}\).  At subleading order the soft
insertion is a differential operator, and multiple soft emissions produce
ordered products and contact terms.  The independent one-photon
multiplication factors used in the leading-log proof are a property of the
leading eikonal term.

The same soft statements have an asymptotic-symmetry formulation.  In
four-dimensional QED, the leading soft photon theorem is equivalent, under
the assumptions on null-infinity limits and charged-state sectors, to a Ward
identity for large \(U(1)\) gauge transformations; its classical counterpart
is electromagnetic memory.  In gravity, the analogous leading and subleading
soft-graviton statements are tied to BMS supertranslations and to the more
delicate superrotation story.  These asymptotic-symmetry interpretations
delimit the leading eikonal approximation used in this chapter; the inclusive
cancellation above uses the exponentiation of \(S^{(0)}\).
\end{remark}

Figure~\ref{fig:volume-ii-soft-factorization} shows the leading eikonal
factorization used in the inclusive calculation: soft photons attach to the
external charged lines, while the hard kernel is unchanged at leading soft
order.

\begin{figure}[!ht]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[>=Stealth,x=1cm,y=1cm,every node/.style={font=\small}]
    \tikzset{
      photon/.style={
        purple!80!black,
        decorate,
        decoration={snake,amplitude=0.75mm,segment length=3.2mm},
        line width=0.95pt,
        -{Latex[length=1.6mm]}
      },
      matter/.style={line width=1.0pt,-{Latex[length=1.8mm]}},
      blob/.style={circle,draw=black,fill=gray!18,line width=0.9pt,minimum size=0.86cm}
    }
    \begin{scope}
      \node[font=\normalsize] at (0,2.3) {one soft photon};
      \node[blob] (B) at (-1.75,0) {};
      \draw[matter] (-3.7,0)--(B.west);
      \draw[matter] (B.east)--(0.35,0);
      \draw[matter] (0.35,0)--(2.4,0);
      \draw[photon] (0.35,0)--(1.18,1.12);
      \node[blue!70!black,below] at (-0.38,-0.12) {\(p+k\)};
      \node[blue!70!black,below] at (1.5,-0.12) {\(p\)};
      \node[purple!80!black,above right] at (1.13,1.05) {\(k,h\)};
      \node[align=center] at (-1.75,-0.92) {hard\\process};
      \node[align=center,blue!65!black] at (0.8,-1.3)
        {\(g\,p\cdot e_h^*/(p\cdot k-\ii\epsilon)\)};
    \end{scope}

    \begin{scope}[shift={(6.2,0)}]
      \node[font=\normalsize] at (0,2.3) {two soft photons on one line};
      \node[blob,minimum size=0.75cm] (C) at (-2.0,0) {};
      \draw[matter] (-3.55,0)--(C.west);
      \draw[matter] (C.east)--(2.5,0);
      \draw[photon] (-0.6,0)--(-0.2,1.05);
      \draw[photon] (0.6,0)--(1.0,1.05);
      \node[purple!80!black] at (-0.12,1.25) {\(k_1,h_1\)};
      \node[purple!80!black] at (1.16,1.25) {\(k_2,h_2\)};
      \node[font=\large] at (2.92,0.06) {\(+\)};
      \begin{scope}[shift={(3.55,0)}]
        \node[blob,minimum size=0.62cm] (D) at (-0.6,0) {};
        \draw[matter] (-1.85,0)--(D.west);
        \draw[matter] (D.east)--(2.0,0);
        \draw[photon] (0.05,0)--(0.42,1.0);
        \draw[photon] (1.08,0)--(1.45,1.0);
      \end{scope}
      \node[align=center,blue!65!black] at (1.5,-1.2)
        {sum over orderings\\equals product of one-photon factors};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{Soft factorization on external charged lines.  A photon with
  wavelength much larger than the hard interaction region resolves only the
  external charge and momentum.}
  \label{fig:volume-ii-soft-factorization}
\end{figure}

\section{Real Soft Emission}

Divide the real-emission contribution by the hard rate with no unresolved
photons.  Using the leading soft factor gives
\[
\begin{aligned}
  R_{\rm real}(E_T,\mu)
  &=
  \sum_{N=0}^{\infty}\frac1{N!}
  \prod_{r=1}^{N}
  \int_{\mu<\omega_r<E_T}
  \frac{\dd^3\vec k_r}{(2\pi)^3\,2\omega_r}
  \sum_{h_r=\pm}
  \left|
    \sum_n
    \frac{g_n\,p_n\cdot e_{h_r}^*(k_r)}
         {\eta_n p_n\cdot k_r}
  \right|^2 .
\end{aligned}
\]
For a single soft photon, define the nonnegative angular kernel
\[
  \mathcal S(k)
  =
  \sum_{h=\pm}
  \left|
    \sum_n
    \frac{g_n\,p_n\cdot e_h^*(k)}
         {\eta_n p_n\cdot k}
  \right|^2 .
\]
A covariant physical-polarization sum may be written as
\[
  \sum_{h=\pm}e_{h,\mu}(k)e^*_{h,\nu}(k)
  =
  \eta_{\mu\nu}+k_\mu c_\nu+k_\nu c_\mu,
  \qquad
  k\cdot c=-1,\qquad c^2=0 .
\]
The terms involving \(c_\mu\) vanish after using
\(\sum_n\eta_n g_n=0\).  Thus
\[
  \mathcal S(k)
  =
  \sum_{n,m}
  \frac{\eta_n\eta_m g_ng_m\,p_n\cdot p_m}
       {(p_n\cdot k)(p_m\cdot k)} .
\]

Writing \(k^\mu=\omega(1,\hat k)\), the energy integral separates from the
angular integral:
\[
  \int_{\mu}^{E_T}
  \frac{\dd^3\vec k}{(2\pi)^3\,2\omega}\,\mathcal S(k)
  =
  \frac{1}{2(2\pi)^3}
  \int_{\mu}^{E_T}\frac{\dd\omega}{\omega}
  \int\dd^2\hat k
  \sum_{n,m}
  \frac{\eta_n\eta_m g_ng_m\,p_n\cdot p_m}
       {(p_n^0-\vec p_n\cdot\hat k)
        (p_m^0-\vec p_m\cdot\hat k)} .
\]
For massive external particles, define the relative-velocity parameter
\[
  \beta_{nm}
  =
  \sqrt{1-\frac{m_n^2m_m^2}{(p_n\cdot p_m)^2}} .
\]
The diagonal term \(n=m\) is understood as the \(\beta_{nm}\to0\) limit.  The
angular integral may be evaluated directly by combining the two denominators:
\[
\begin{aligned}
  &\int\dd^2\hat k\,
  \frac{1}
       {(p_n^0-\vec p_n\cdot\hat k)
        (p_m^0-\vec p_m\cdot\hat k)}
\\
  &=
  \int_0^1\dd x
  \int\dd^2\hat k\,
  \frac{1}
  {\left[
    p_n^0x+p_m^0(1-x)
    -(\vec p_nx+\vec p_m(1-x))\cdot\hat k
  \right]^2}.
\end{aligned}
\]
Choose the polar axis along
\(\vec p_nx+\vec p_m(1-x)\).  The \(\hat k\)-integral becomes an elementary
integral of
\[
  \frac{2\pi\sin\theta\,\dd\theta}
  {\left[
    p_n^0x+p_m^0(1-x)
    -|\vec p_nx+\vec p_m(1-x)|\cos\theta
  \right]^2}.
\]
Performing the \(x\)-integral gives
\[
  \int\dd^2\hat k\,
  \frac{1}
       {(p_n^0-\vec p_n\cdot\hat k)
        (p_m^0-\vec p_m\cdot\hat k)}
  =
  -\frac{2\pi}{p_n\cdot p_m}\,
  \frac{1}{\beta_{nm}}
  \log\frac{1+\beta_{nm}}{1-\beta_{nm}} .
\]
It is useful to introduce
\[
  A_{\beta\alpha}
  =
  -\frac{1}{8\pi^2}
  \sum_{n,m}
  \frac{\eta_n\eta_m g_ng_m}{\beta_{nm}}
  \log\frac{1+\beta_{nm}}{1-\beta_{nm}} .
\]
The representation by \(\mathcal S(k)\) shows \(A_{\beta\alpha}\ge0\) for the
massive charged scattering considered here.  The real unresolved photons
therefore give
\[
  R_{\rm real}(E_T,\mu)
  =
  \sum_{N=0}^{\infty}
  \frac1{N!}
  \left[
    A_{\beta\alpha}\log\frac{E_T}{\mu}
  \right]^N
  =
  \left(\frac{E_T}{\mu}\right)^{A_{\beta\alpha}}
\]
at leading soft logarithmic order.

\section{Virtual Soft Photons and External Factors}

A virtual soft photon exchanged between two distinct external charged lines
also factorizes.  For external lines \(n\) and \(m\), set
\[
  J_{nm}
  =
  -\ii\,p_n\cdot p_m
  \int_{\mu<|\vec k|<M}
  \frac{\dd^4k}{(2\pi)^4}\,
  \frac{1}
       {(k^2-\ii\epsilon)
        (\eta_n p_n\cdot k-\ii\epsilon)
        (-\eta_m p_m\cdot k-\ii\epsilon)} .
\]
The virtual soft photons exchanged among distinct external lines exponentiate:
\[
  \exp\!\left[
    \frac12\sum_{n\ne m}g_ng_mJ_{nm}
  \right].
\]
This expression omits the soft photons whose two endpoints lie on the same
external charged line.  Those terms are tied to the infrared part of the LSZ
normalization factors.  A direct way to see the accounting is to insert the
electromagnetic current
\[
  \widehat j^\mu(x)=\partial_\nu \widehat F^{\mu\nu}(x)
\]
between one-particle states.  In scalar QED, with
\(k=p_2-p_1\) and
\(\omega_p=\sqrt{\vec p^{\,2}+m^2}\),
\[
  \langle \vec p_2|\widehat j^\mu(0)|\vec p_1\rangle
  =
  \frac{g}{(2\pi)^3\sqrt{2\omega_{p_1}}\sqrt{2\omega_{p_2}}}
  (p_1+p_2)^\mu F(k^2),
  \qquad F(0)=1 .
\]
In spinor QED the corresponding decomposition is
\[
\begin{aligned}
  &\langle \vec p_2,\sigma_2|\widehat j^\mu(0)
  |\vec p_1,\sigma_1\rangle
\\
  &\qquad =
  \frac{\ii g}{(2\pi)^3}\,
  \bar u^{\sigma_2}(\vec p_2)
  \left[
    \gamma^\mu F(k^2)
    -\frac{\ii}{2m}(p_1+p_2)^\mu G(k^2)
  \right]
  u^{\sigma_1}(\vec p_1),
  \qquad F(0)+G(0)=1 .
\end{aligned}
\]
In these conventions the one-loop value \(G(0)=-g^2/(8\pi^2)+O(g^4)\) is a
finite contribution to the magnetic form factor.  The electromagnetic matrix
element at \(k^2=0\) is free of infrared divergence, although derivatives such
as \(\left.\dd F/\dd k^2\right|_{k^2=0}\) need not be.  Thus the soft virtual
photons attached to the current insertion must be accompanied by the same
infrared part of the LSZ factors.
In the current form factor, soft photons connecting
the two charged legs exponentiate as \(e^{g^2J_{12}}\).  Finiteness at
\(k^2=0\) then fixes the infrared factor in \(Z\):
\[
  Z_{\rm IR}
  =
  \left.
  e^{-g^2J_{12}}
  \right|_{p_2=p_1,\ \eta_2=-\eta_1}.
\]

Equivalently, compare this with the same-line expression obtained by setting
the two eikonal velocities equal before interpreting the self-contraction:
\[
  ``J_{11}''
  =
  -\ii p_1^2
  \int_{\mu<|\vec q|<M}
  \frac{\dd^4q}{(2\pi)^4}
  \frac{1}
       {(q^2-\ii\epsilon)
        (p_1\cdot q-\ii\epsilon)
        (-p_1\cdot q-\ii\epsilon)} .
\]
The two matter poles in the \(q^0\)-plane contribute to the imaginary part.
The real part satisfies
\[
  \operatorname{Re}J_{11}
  =
  -\left.
  \operatorname{Re}J_{12}
  \right|_{p_2=p_1,\ \eta_2=-\eta_1},
  \qquad
  Z_{\rm IR}=e^{g^2\operatorname{Re}J_{11}} .
\]

At one-loop order this same statement follows directly from the scalar-QED
self-energy.  The soft part is
\[
  \ii\Sigma(p)
  =
  \int\frac{\dd^4k}{(2\pi)^4}
  \frac{-\ii}{(p+k)^2+m^2-\ii\epsilon}
  \frac{-\ii}{k^2-\ii\epsilon}
  (\ii g)^2(2p+k)^2+\cdots ,
\]
whose infrared approximation is
\[
  \ii\Sigma(p)\big|_{\rm soft}
  =
  4g^2p^2
  \int_{|\vec k|<M}
  \frac{\dd^4k}{(2\pi)^4}
  \frac{1}{p^2+m^2+2p\cdot k-\ii\epsilon}
  \frac{1}{k^2-\ii\epsilon}.
\]
The corresponding external factor
\[
  Z
  =
  \left(
    1-\left.
    \frac{\partial\Sigma}{\partial p^2}\right|_{p^2=-m^2}
  \right)^{-1}
\]
contains the same soft logarithm as the same-line eikonal loop.  In the
infrared region,
\[
  \log Z_{\rm IR}
  =
  \ii g^2\,4p^2
  \int_{|\vec k|<M}
  \frac{\dd^4k}{(2\pi)^4}
  \frac{1}
       {(2p\cdot k-\ii\epsilon)^2(k^2-\ii\epsilon)}
  =
  g^2\operatorname{Re}J_{11}+O(g^4).
\]
Thus soft loops beginning and ending on a single external charged line are
not additional uncancelled infrared factors; they are already included in
the LSZ normalization of that charged external line.

Figure~\ref{fig:volume-ii-same-line-ir-cancellation} identifies the same-line
soft loop contribution and its placement in the external-line normalization
factor rather than as a separate inclusive-rate correction.

\begin{figure}[!ht]
  \centering
  \resizebox{0.82\textwidth}{!}{%
  \begin{tikzpicture}[>=Stealth,x=1cm,y=1cm,every node/.style={font=\small}]
    \tikzset{
      photon/.style={purple!80!black,decorate,
        decoration={snake,amplitude=0.72mm,segment length=3.1mm},
        line width=0.95pt},
      matter/.style={line width=1.0pt,-{Latex[length=1.8mm]}},
      blob/.style={circle,draw=black,fill=gray!18,line width=0.9pt,minimum size=0.75cm},
      zblob/.style={circle,draw=purple!80!black,fill=purple!12,line width=0.9pt,
        minimum size=0.42cm}
    }
    \begin{scope}
      \node[font=\normalsize] at (0,2.05) {current form factor};
      \draw[matter] (-3.3,0)--(-0.45,0);
      \node[blob] at (0,0) {};
      \draw[matter] (0.45,0)--(3.3,0);
      \draw[photon] (0,-1.25)--(0,-0.42);
      \draw[photon] (-2.05,0) .. controls (-1.55,1.05) and
        (1.25,1.05) .. (1.85,0);
      \node[purple!80!black] at (-0.1,1.28) {soft virtual photons};
      \node[blue!70!black] at (-2.65,-0.35) {\(Z^{1/2}\)};
      \node[blue!70!black] at (2.65,-0.35) {\(Z^{1/2}\)};
      \node[purple!80!black,below] at (0,-1.25) {\(j^\mu\)};
    \end{scope}
    \begin{scope}[shift={(7.3,0)}]
      \node[font=\normalsize] at (0,2.05) {same-line soft loops};
      \draw[matter] (-3.25,0)--(3.25,0);
      \node[zblob] at (-1.65,0) {};
      \node[zblob] at (-0.55,0) {};
      \draw[photon] (-0.55,0) .. controls (-0.15,0.85) and
        (1.3,0.85) .. (1.7,0);
      \draw[photon] (0.2,0) .. controls (0.45,0.45) and
        (1.0,0.45) .. (1.25,0);
      \node[purple!80!black] at (-1.1,0.62) {\(Z_{\rm IR}\)};
      \node[purple!80!black] at (1.35,1.1) {absorbed into \(Z_{\rm IR}\)};
      \node[align=center,blue!70!black] at (0,-0.78)
        {no separate same-line\\infrared factor remains};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The infrared part of the external LSZ factor accounts for soft
  virtual photons whose two endpoints lie on the same charged external line.}
  \label{fig:volume-ii-same-line-ir-cancellation}
\end{figure}

Combining the distinct-line soft loops with the infrared parts of the
\(Z_n^{1/2}\) factors gives
\[
  \exp\!\left[
    \frac12\sum_{n\ne m}g_ng_mJ_{nm}
  \right]\prod_n (Z_n^{\rm IR})^{1/2}
  =
  (\hbox{phase})\,
  \exp\!\left[
    \frac12\sum_{n,m}g_ng_m\,\operatorname{Re}J_{nm}
  \right].
\]

The \(k^0\)-plane poles of \(J_{nm}\), including only the denominators that
control the contour placement, are
\[
  -|\vec k|+\ii\epsilon,\qquad
  |\vec k|-\ii\epsilon,\qquad
  \frac{\vec p_n\cdot\vec k}{p_n^0}-\ii\eta_n\epsilon,\qquad
  \frac{\vec p_m\cdot\vec k}{p_m^0}+\ii\eta_m\epsilon .
\]
The photon poles give the real part of the eikonal integral, while the
matter poles may contribute to its phase.  Evaluating the photon-pole
contribution gives
\[
  \operatorname{Re}J_{nm}
  =
  -\eta_n\eta_m\,
  \frac{p_n\cdot p_m}{2}
  \int_{\mu<|\vec k|<M}
  \frac{\dd^3\vec k}{(2\pi)^3}
  \frac{1}
       {|\vec k|^3
        (p_n^0-\vec p_n\cdot\hat k)
        (p_m^0-\vec p_m\cdot\hat k)} .
\]
Using the same angular integral as for real emission,
\[
  \operatorname{Re}J_{nm}
  =
  \eta_n\eta_m\,
  \frac{1}{8\pi^2\beta_{nm}}
  \log\frac{1+\beta_{nm}}{1-\beta_{nm}}
  \log\frac{M}{\mu}.
\]
Therefore the virtual factor in the amplitude has modulus
\[
  \left(\frac{\mu}{M}\right)^{A_{\beta\alpha}/2},
\]
and its contribution to a rate is
\[
  R_{\rm virt}(M,\mu)
  =
  \left(\frac{\mu}{M}\right)^{A_{\beta\alpha}} .
\]

Figure~\ref{fig:volume-ii-bloch-nordsieck-cancellation} displays the
Bloch--Nordsieck cancellation in the notation above: the virtual factor and the
real-emission inclusive factor carry opposite powers of the soft regulator.

\begin{figure}[!ht]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[>=Stealth,x=1cm,y=1cm,every node/.style={font=\small}]
    \tikzset{
      photon/.style={purple!80!black,decorate,
        decoration={snake,amplitude=0.72mm,segment length=3.1mm},
        line width=0.95pt},
      matter/.style={line width=1.0pt,-{Latex[length=1.8mm]}},
      blob/.style={circle,draw=black,fill=gray!18,line width=0.9pt,minimum size=0.78cm}
    }
    \begin{scope}
      \node[font=\normalsize] at (0,2.45) {real unresolved photons};
      \node[blob] (A) at (-1.6,0.2) {};
      \draw[line width=1.0pt] (-3.1,0.95)--(A);
      \draw[line width=1.0pt] (-3.1,-0.45)--(A);
      \draw[matter] (A)--(1.35,1.1);
      \draw[matter] (A)--(1.35,-0.68);
      \foreach \x/\y/\lab in {-0.55/0.18/{k_1},0.1/0.02/{k_2},0.63/0.45/{k_N}} {
        \draw[photon,-{Latex[length=1.45mm]}] (\x,\y)--(\x+0.42,\y+0.65);
        \node[purple!80!black] at (\x+0.57,\y+0.75) {\(\lab\)};
      }
      \node[align=center,blue!65!black] at (-0.25,-1.35)
        {\(\displaystyle R_{\rm real}
        =\left(E_T/\mu\right)^{A_{\beta\alpha}}\)};
    \end{scope}

    \begin{scope}[shift={(5.2,0)}]
      \node[font=\normalsize] at (0,2.45) {virtual soft cloud};
      \draw[line width=1.0pt] (-2.35,1.0)--(2.35,1.22);
      \draw[line width=1.0pt] (-2.35,-0.72)--(2.35,-0.94);
      \foreach \x in {-1.25,-0.28,0.66} {
        \draw[photon] (\x,1.03)--(\x+0.34,-0.82);
      }
      \draw[photon] (1.38,1.16) .. controls (2.05,0.45) and (1.93,-0.12) .. (1.55,-0.92);
      \node[align=center] at (0,-1.52)
        {soft loops plus \(Z_n^{1/2}\)};
      \node[align=center,blue!65!black] at (0,-2.28)
        {\(\displaystyle R_{\rm virt}
        =\left(\mu/M\right)^{A_{\beta\alpha}}\)};
    \end{scope}

    \begin{scope}[shift={(10.4,0)}]
      \node[font=\normalsize] at (0,2.45) {inclusive rate};
      \draw[rounded corners=2pt,fill=green!8,draw=green!45!black,line width=0.9pt]
        (-2.15,-0.85) rectangle (2.15,1.15);
      \node[align=center] at (0,0.73)
        {same observed hard event\\for photons below \(E_T\)};
      \draw[photon] (-1.05,-0.55)--(-0.48,-0.05);
      \draw[photon] (0.18,-0.58)--(0.82,-0.12);
      \node[align=center,blue!65!black] at (0,-1.5)
        {\(\displaystyle
        \left(\mu/M\right)^{A_{\beta\alpha}}
        \left(E_T/\mu\right)^{A_{\beta\alpha}}
        =
        \left(E_T/M\right)^{A_{\beta\alpha}}\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{Real unresolved photons and virtual soft photons carry opposite
  powers of the auxiliary infrared cutoff.  The product is finite for a fixed
  detector threshold.}
  \label{fig:volume-ii-bloch-nordsieck-cancellation}
\end{figure}

\section{Inclusive Probabilities}
\label{sec:volume-ii-inclusive-probabilities}

The cancellation of the auxiliary photon-mass regulator between real
unresolved emission and virtual soft exchange is the Bloch--Nordsieck
cancellation.  Equivalently, real soft emission cancels virtual infrared
divergences at the level of inclusive probabilities.  The order of operations
is part of the statement.  Fix
\[
  0<\mu<E_T<M
\]
and compute in the regulated theory.  At fixed \(\mu\), first include all
detector-degenerate unresolved photons, or, coefficientwise in perturbation
theory, all detector-degenerate terms contributing to the chosen coefficient.
Only after that inclusive sum has been formed is the regulator
\(\mu\downarrow0\) removed.

In the leading soft approximation this ordered construction is completely
explicit.  Put
\[
  \ell_T(\mu)=\log\frac{E_T}{\mu},
  \qquad
  \ell_M(\mu)=\log\frac{M}{\mu}.
\]
The real contribution with exactly \(N\) unresolved photons is
\[
  R_N(E_T,\mu)
  =
  \frac{1}{N!}
  \left[A_{\beta\alpha}\ell_T(\mu)\right]^N ,
\]
and the virtual soft factor in the rate is
\[
  R_{\rm virt}(M,\mu)
  =
  \exp\{-A_{\beta\alpha}\ell_M(\mu)\}.
\]
Thus the leading-soft regulated inclusive factor is
\[
  R_{\rm incl}^{(\mu)}(E_T,M)
  =
  R_{\rm virt}(M,\mu)
  \sum_{N=0}^\infty R_N(E_T,\mu).
\]
The infinite sum over \(N\) is taken at fixed positive \(\mu\), where it is an
ordinary exponential series.  It gives
\[
  R_{\rm incl}^{(\mu)}(E_T,M)
  =
  \exp\{-A_{\beta\alpha}\ell_M(\mu)\}
  \exp\{A_{\beta\alpha}\ell_T(\mu)\}
  =
  \left(\frac{E_T}{M}\right)^{A_{\beta\alpha}},
\]
which is independent of \(\mu\).  Multiplying by the hard finite factor gives
the leading soft result
\[
  P_{\beta|\alpha}(E_T)
  =
  P_{\beta|\alpha}^{\rm hard}(M)\,
  \left(\frac{E_T}{M}\right)^{A_{\beta\alpha}}
  \times
  F_{\beta\alpha}(E_T/M),
\]
where \(F_{\beta\alpha}\) is finite as \(\mu\to0\) and equals \(1\) in the
leading logarithmic approximation.  The scale \(M\) separates hard and soft
physics; dependence on it cancels once the hard finite correction is computed
to the same order.

\begin{proposition}[Coefficientwise Bloch--Nordsieck cancellation]
\label{prop:coefficientwise-bloch-nordsieck}
Let \(A_{\beta\alpha}\) be regarded as a formal quantity of order \(g^2\).
At every fixed power \(A_{\beta\alpha}^L\), the coefficient of the inclusive
leading-soft rate has a regulator-independent limit after the sum over
unresolved photon numbers contributing to that coefficient is taken.  More
precisely,
\[
\begin{aligned}
  &[A_{\beta\alpha}^L]\,
  \left[
    \exp\{-A_{\beta\alpha}\ell_M(\mu)\}
    \sum_{N=0}^{\infty}
    \frac{\left[A_{\beta\alpha}\ell_T(\mu)\right]^N}{N!}
  \right]
\\
  &\hspace{2cm}=
  \sum_{N=0}^{L}
  \frac{[-\ell_M(\mu)]^{L-N}}{(L-N)!}
  \frac{\ell_T(\mu)^N}{N!}
  =
  \frac{1}{L!}
  \left(\ell_T(\mu)-\ell_M(\mu)\right)^L
\\
  &\hspace{2cm}=
  \frac{1}{L!}
  \left(\log\frac{E_T}{M}\right)^L .
\end{aligned}
\]
Thus the cancellation is not a statement about an individual fixed photon
number.  It is a statement about the inclusive coefficient obtained after the
degenerate real-emission terms and the virtual terms at the same perturbative
order are combined.
\end{proposition}

\begin{proof}
At fixed \(\mu>0\), the product of the virtual exponential and the real
emission exponential is an identity of formal power series in
\(A_{\beta\alpha}\).  The coefficient of \(A_{\beta\alpha}^L\) receives
contributions from \(N\) real photons and \(L-N\) powers of the virtual
exponential, with \(0\le N\le L\).  This gives the displayed binomial sum.
Since
\[
  \ell_T(\mu)-\ell_M(\mu)
  =
  \log\frac{E_T}{\mu}-\log\frac{M}{\mu}
  =
  \log\frac{E_T}{M},
\]
all dependence on the auxiliary regulator cancels before the limit
\(\mu\downarrow0\) is taken.
\end{proof}

For a fixed finite number \(N\) of photons, by contrast, the rate contains
\[
  \left(\frac{\mu}{M}\right)^{A_{\beta\alpha}}
  \frac1{N!}
  \left[
    A_{\beta\alpha}\log\frac{E_T}{\mu}
  \right]^N ,
\]
which tends to zero as \(\mu\to0\) when \(A_{\beta\alpha}>0\).  The
resolution-inclusive sum remains finite because arbitrarily many unresolved
soft photons are included.

The coupling expansion and the infrared limit are therefore nonuniform in the
exclusive fixed-\(N\) decomposition.  The perturbative statement is the
coefficientwise one in
Proposition~\ref{prop:coefficientwise-bloch-nordsieck}: coefficient extraction
is performed at fixed regulator, all degenerate real and virtual terms at that
coefficient are added, and only then is \(\mu\) removed.  The exponentiated
leading-soft expression is a resummation of these coefficientwise
cancellations, not a license to take the regulator limit term by term in
photon number.

\begin{remark}
The fixed-photon-number matrix elements are exclusive kernels for an idealized
measurement that specifies the exact number of photons down to zero energy.
The inclusive probability above is the transition probability associated with
a finite energy resolution.  In a charged sector of massless QED, the
exclusive kernels should not be reinterpreted as ordinary regulator-independent
LSZ \(S\)-matrix elements for bare electron states.  Both the vanishing of
fixed-\(N\) probabilities and the finiteness of the inclusive probability
follow from the same soft factorization formulas.
\end{remark}

The discussion so far treated massive charged particles, where the angular
integral is finite.  If massless charged particles occur, photon emission
collinear to a charged external line produces additional singularities.  The
soft Bloch--Nordsieck sum is then only part of the required inclusive
construction.  The finite-degeneracy statement behind the
Kinoshita--Lee--Nauenberg prescription is the following.

\begin{theorem}[Finite-degeneracy KLN cancellation]
\label{thm:finite-degeneracy-kln}
Let \(\mathcal H_0\) be a Hilbert space with unperturbed Hamiltonian \(H_0\).
Let \(P_A\) and \(P_B\) be the orthogonal projectors onto two
finite-dimensional degenerate subspaces of \(H_0\), with
\[
  d_A=\operatorname{rank}P_A<\infty,\qquad
  d_B=\operatorname{rank}P_B<\infty .
\]
Let \(H=H_0+V\), where \(V\) is a bounded self-adjoint perturbation, and let
\[
  U_T
  =
  \mathcal T\exp\left\{-\ii\int_{-T}^{T}
    e^{\ii H_0t}V e^{-\ii H_0t}\,\dd t\right\}
\]
be the interaction-picture evolution, possibly with an adiabatic switch-on
understood.  Assume that, after introducing whatever ultraviolet and infrared
regulators are being used, the inclusive long-time limit
\[
  \mathcal P^{(\rho)}_{B|A}
  =
  \lim_{T\to\infty}
  \frac{1}{d_A}\operatorname{Tr}_{\mathcal H_0}
  \left(P_B U_T P_A U_T^\dagger P_B\right)
\]
exists coefficientwise in the perturbation expansion.  Finally assume that
the regulator-singular factors associated with exact degeneracy act by
unitary rotations inside the degenerate subspaces: for the
\(P_A\to P_B\) block of the regulated scattering operator,
\[
  P_B S_\rho P_A
  =
  U_B(\rho)\,S_{\rm fin}(\rho)\,U_A(\rho)^\dagger ,
\]
where \(U_A(\rho)\) and \(U_B(\rho)\) are unitary on \(P_A\mathcal H_0\) and
\(P_B\mathcal H_0\), respectively, and \(S_{\rm fin}(\rho)\) has a finite
coefficientwise limit as the regulator \(\rho\downarrow0\).  Then
\[
  \lim_{\rho\downarrow0}\mathcal P^{(\rho)}_{B|A}
  =
  \frac{1}{d_A}
  \operatorname{Tr}_{P_B\mathcal H_0}
  \left(S_{\rm fin}S_{\rm fin}^\dagger\right)
\]
exists coefficientwise.  The final-state sum removes the outgoing
degenerate rotation \(U_B(\rho)\), and the initial-state average removes the
incoming degenerate rotation \(U_A(\rho)\).
\end{theorem}

\begin{proof}
The normalized incoming density matrix for an unresolved initial degeneracy
class is
\[
  \varrho_A=\frac{P_A}{d_A}.
\]
The inclusive probability of finding the outgoing state in \(P_B\mathcal H_0\)
is
\[
  \operatorname{Tr}_{\mathcal H_0}
  \left(P_B S_\rho \varrho_A S_\rho^\dagger P_B\right)
  =
  \frac1{d_A}
  \operatorname{Tr}_{P_B\mathcal H_0}
  \left(
    P_B S_\rho P_A S_\rho^\dagger P_B
  \right).
\]
Insert
\[
  P_B S_\rho P_A
  =
  U_B(\rho)S_{\rm fin}(\rho)U_A(\rho)^\dagger .
\]
Because \(U_A(\rho)^\dagger U_A(\rho)=P_A\) on the incoming degenerate
subspace and \(U_B(\rho)^\dagger U_B(\rho)=P_B\) on the outgoing degenerate
subspace, cyclicity of the finite-dimensional trace gives
\[
\begin{aligned}
  \mathcal P^{(\rho)}_{B|A}
  &=
  \frac1{d_A}
  \operatorname{Tr}_{P_B\mathcal H_0}
  \left[
    U_B(\rho)S_{\rm fin}(\rho)
    S_{\rm fin}(\rho)^\dagger U_B(\rho)^\dagger
  \right]
\\
  &=
  \frac1{d_A}
  \operatorname{Tr}_{P_B\mathcal H_0}
  \left[
    S_{\rm fin}(\rho)S_{\rm fin}(\rho)^\dagger
  \right].
\end{aligned}
\]
The assumed coefficientwise limit of \(S_{\rm fin}(\rho)\) therefore gives the
claimed coefficientwise limit of the inclusive probability.
\end{proof}

The theorem is an abstract finite-regulator statement.  In a relativistic
field theory the interaction is generally an unbounded operator-valued
distribution on the physical Hilbert space, and exact massless degeneracy is
typically continuous rather than finite-dimensional.  The theorem identifies
the finite-volume, energy-resolution, and regulator structure that must be
implemented before the regulator limits are taken.  After those
regularizations have turned detector-degenerate sectors into finite
subspaces, the inclusive construction sums over final states and averages over
initial states inside the same detector degeneracy class.  In symbols, if
\([\alpha]\) and \([\beta]\) are the incoming and outgoing equivalence classes
defined by the detector resolution, one considers
\[
  P([\alpha]\to[\beta])
  =
  \sum_{\alpha'\in[\alpha]}\omega_{\alpha'}
  \sum_{\beta'\in[\beta]}
  \left|S^{(\mu)}_{\beta'\alpha'}\right|^2,
\]
with nonnegative weights \(\omega_{\alpha'}\) normalized on the incoming
degenerate class.  Soft degeneracy and collinear degeneracy are therefore
handled by the same operational rule: the transition probability is attached
to what the detector distinguishes.

\begin{remark}[Abelian scope of Bloch--Nordsieck cancellation]
\label{rem:abelian-scope-bloch-nordsieck}
The Bloch--Nordsieck exponentiation derived in this chapter is an Abelian
soft-photon statement.  Its leading real-emission factor is a scalar function
of the hard charges and momenta, and different soft photons multiply by
commuting factors.  Nonabelian gauge theory changes this algebraic input:
soft gluons couple through color generators, and the soft factors are
operators in color space, naturally organized by Wilson lines and color
ordered products.

Consequently, the QED formula above should not be transplanted directly to
partonic QCD.  In perturbative QCD, final-state soft and collinear
singularities cancel in sufficiently inclusive color-singlet observables or
factor into universal jet and soft functions, but initial-state collinear
singularities of colored partons are absorbed into the renormalization of
parton distributions.  The finite hadronic observable is then obtained from
renormalized light-ray matrix elements and short-distance kernels, as in the
DIS factorization formula
\eqref{eq:dis-factorization} and DGLAP equation
\eqref{eq:dglap-evolution} of
Chapter~\ref{ch:volume-ii-19}.  This is the precise replacement, in QCD
factorization, for an abelian fixed-charge Bloch--Nordsieck cancellation.
\end{remark}

\section{Soft Photon Number Distribution}

The leading soft formula also determines the unresolved photon multiplicity
distribution.  Let
\[
  \overline N_{\beta\alpha}(\mu,E_T)
  =
  A_{\beta\alpha}\log\frac{E_T}{\mu}
\]
be the mean number of photons in the energy window
\(\mu<\omega<E_T\) generated by the eikonal current.  The probability for
emitting exactly \(N\) photons in this unresolved window has the Poisson form
\[
  P_N(\mu,E_T)
  =
  \exp\{-\overline N_{\beta\alpha}\}
  \frac{\overline N_{\beta\alpha}^{\,N}}{N!}
\]
at leading soft order, multiplied by the hard probability and finite
corrections.  The exponential factor is the virtual soft factor together with
the no-resolved-emission probability.  Summing over \(N\) gives unity for the
soft probability distribution, while the hard probability is multiplied by the
finite detector-resolution factor
\[
  \left(\frac{E_T}{M}\right)^{A_{\beta\alpha}} .
\]

As \(\mu\to0\), the mean \(\overline N_{\beta\alpha}\) diverges when
\(A_{\beta\alpha}>0\).  The inclusive probability is finite because the
detector state is a distribution over arbitrarily many unresolved photons.
The exclusive fixed-\(N\) probability tends to zero because it samples a set
of vanishing measure inside the infrared cloud selected by the long-range
electromagnetic field.

\section{Charged Asymptotic Sectors and Infraparticles}

The preceding formulas describe a scattering problem whose charged asymptotic
representation is fixed by Gauss' law.  Let \(\mathfrak A_{\mathrm{loc}}\) denote
the algebra of gauge-invariant observables localized in bounded spacetime
regions, and let
\[
  E^i=F^{i0}
\]
be the electric field.  At a fixed time, for a smooth function \(f\) on the
unit sphere, define the large-radius electric-flux observable formally by
\[
  \Phi_R(f)
  =
  \int_{S^2}\dd\Omega(\mathbf n)\,
  R^2 f(\mathbf n)\,n_i E^i(0,R\mathbf n).
\]
The constant test function gives the total electric charge in the limit
\[
  Q=\lim_{R\to\infty}\Phi_R(1),
\]
when the limit exists in the representation under consideration.  More
generally, the limiting angular flux functional
\[
  \Phi_\infty(f)=\lim_{R\to\infty}\Phi_R(f)
\]
records the electromagnetic field at spacelike infinity.  Local observables
cannot change this limiting flux: if \(A\in\mathfrak A_{\mathrm{loc}}\) has
support in a fixed bounded region, then for \(R\) large enough the flux
surface is spacelike separated from the support of \(A\), and locality makes
the commutator vanish before the limit is taken.  Thus the limiting flux, when
it is represented sharply, is a superselection datum for the charged sector.
For \(f=1\) this is the usual electric charge superselection rule; for
nonconstant \(f\) it is the angular Coulomb field attached to the asymptotic
charged motion.

There are two logically separate consequences.  First, a bounded local
gauge-invariant observable cannot change the limiting surface charge: for a
large sphere spacelike to its localization region it commutes with the
surface-flux operator before the \(R\to\infty\) limit.  This is the
localization obstruction to creating a charged particle by ordinary local
observable Haag--Ruelle operators.  Second, in massless QED the charged sector
also carries velocity-dependent soft electromagnetic flux, so its spectral
support is not an ordinary sharp Wigner mass shell.  The next theorem concerns
this second statement.  The Wilson-line and Coulomb dressings needed for the
first statement are formulated as a scattering-theory problem in
Chapter~\ref{ch:volume-iv-haag-ruelle-mathematical-scattering}.

\begin{theorem}[Buchholz infraparticle obstruction]
\label{thm:buchholz-infraparticle-obstruction}
In a local relativistic quantum field theory satisfying Maxwell's equations
and Gauss' law, a pure state carrying nonzero abelian gauge charge cannot be a
Wigner one-particle state with a sharp mass shell.  More precisely, charged
particle weights are accompanied by asymptotic electromagnetic flux, and the
representations induced by different charged velocities need not be unitarily
equivalent.  In particular, the charged sector is not an ordinary tensor
product of a Wigner electron one-particle representation with the vacuum
photon Fock representation, and Lorentz boosts of a charged particle change
the flux sector rather than acting inside one fixed Fock sector.
\end{theorem}

The electron in four-dimensional QED is therefore an infraparticle.  The
conclusion is an algebraic consequence of locality and Gauss'
law.\footnote{The theorem-level result used here is Buchholz's Gauss-law
analysis of the infraparticle problem.  The spectral analysis of charged
infraparticles was developed by Fr\"ohlich--Morchio--Strocchi, while the
coherent asymptotic construction used in perturbation theory is the
Faddeev--Kulish construction.  The review of Buchholz--Dybalski gives a
modern scattering-theoretic account of these relations.}  Perturbation theory
displays the same point in a direct way.
For a charged particle of velocity
\[
  \mathbf v=\frac{\mathbf p}{p^0},
\]
the Coulomb field boosted to that velocity has radial flux density
\begin{equation}
  \mathcal E_{g,\mathbf v}(\mathbf n)
  =
  \frac{g}{4\pi}
  \frac{1-|\mathbf v|^2}
       {(1-\mathbf v\cdot\mathbf n)^2},
  \qquad
  \int_{S^2}\dd\Omega(\mathbf n)\,
  \mathcal E_{g,\mathbf v}(\mathbf n)=g .
  \label{eq:boosted-coulomb-flux-density}
\end{equation}
Different velocities give different angular flux functionals.  A multiparticle
charged configuration carries the sum of the corresponding flux densities.
This flux data is what the soft photon cloud implements.

The non-Fock nature of the dressing can already be read from the one-photon
Hilbert norm.  The physical photon one-particle space has norm
\[
  \|f\|_{\gamma}^2
  =
  \sum_h
  \int
  \frac{\dd^3\mathbf k}{(2\pi)^3\,2\omega}\,
  |f^h(k)|^2 .
\]
For the leading charged soft profile
\[
  f_p^h(k)\sim g\,\frac{p\cdot e_h^*(k)}{p\cdot k},
  \qquad \omega=|\mathbf k|,
\]
one has \(p\cdot k=\omega(p^0-\mathbf p\cdot\mathbf n)\) and therefore
\[
  \|f_p\|_{\gamma}^2
  \sim
  C(p,g)\int_0^\Lambda\frac{\dd\omega}{\omega},
  \qquad C(p,g)>0\quad(g\ne0).
\]
Thus \(f_p\notin\mathcal H_{1,\gamma}\) after the infrared cutoff is removed.
At finite cutoff the coherent operator below is a unitary Weyl operator on
photon Fock space; in the limit it defines an inequivalent representation
labelled by the asymptotic flux.  If two dressings have different limiting
flux, their difference is not square-integrable in the same norm, so they are
assigned to inequivalent infrared representations rather than to two vectors
in one common photon Fock representation.  The inclusive construction avoids
choosing a sharp charged flux sector by summing over detector-unresolved
radiation; the dressed construction chooses a sector adapted to the hard
charged asymptotic data.

\begin{openproblem}[Charged asymptotic completeness in massless QED]
\label{op:qed-charged-asymptotic-completeness}
Give a nonperturbative Hilbert-space construction of charged scattering in
four-dimensional QED with the following data and conclusions.  For each
charged superselection sector \(\mathcal H_{\rm phys}^{\rm ch}\), construct an
asymptotic space
\[
  \mathcal H_{\rm as}^{\rm ch}
\]
whose labels include hard charged velocity and spin data, ordinary transverse
photon radiation, and the limiting angular electric-flux data required by
Gauss' law.  Construct large-time wave operators
\[
  \Omega_{\rm in/out}^{\rm ch}:
  \mathcal H_{\rm as}^{\rm ch}
  \longrightarrow
  \mathcal H_{\rm phys}^{\rm ch}
\]
which intertwine spacetime translations with the total asymptotic
energy-momentum, represent the limiting electromagnetic flux consistently with
\(\Phi_\infty\), and are isometric after the appropriate infrared
identifications of zero-resolution photon clouds.  The asymptotic-completeness
claim is the equality
\[
  \overline{\operatorname{Ran}\Omega_{\rm in}^{\rm ch}}
  =
  \overline{\operatorname{Ran}\Omega_{\rm out}^{\rm ch}}
  =
  \mathcal H_{\rm phys}^{\rm ch}
\]
in the chosen charged sector, together with a scattering operator on
\(\mathcal H_{\rm as}^{\rm ch}\) whose detector-inclusive probabilities agree
with the regulator-independent inclusive probabilities constructed above from
Bloch--Nordsieck and KLN degeneracy sums.

The theorem above identifies the obstruction that fixes the form of the
problem: the asymptotic space cannot be the tensor product of a Wigner
one-electron Fock space with the vacuum photon Fock space.  A charged creator
must carry electromagnetic flux to infinity; in a gauge-invariant description
this requires a noncompact dressing, for example a Dirac-Coulomb dressing or a
Wilson-line dressing, and a corresponding replacement for the ordinary
local-field LSZ limit.  The Faddeev--Kulish construction supplies the
perturbative coherent soft datum suggested by the leading eikonal factor, and
the inclusive construction supplies finite detector probabilities.  The open
problem is to put these structures into a single nonperturbative large-time
scattering theorem.
\end{openproblem}

\section{Dressed Charged States}

The same eikonal data can also be incorporated into the asymptotic states.
For a charged particle of momentum \(p\), a soft coherent dressing has the
leading soft form
\[
  \mathcal W_p
  =
  \exp\!\left[
    \sum_h
    \int_{\omega<\Lambda}
    \frac{\dd^3\vec k}{(2\pi)^3\,2\omega}
    \left(
      f_p^h(k)a_h^\dagger(k)
      -
      \overline{f_p^h(k)}a_h(k)
    \right)
  \right],
\]
with leading coefficient
\[
  f_p^h(k)\sim g\,\frac{p\cdot e_h^*(k)}{p\cdot k}.
\]
The operator \(\mathcal W_p\) attaches the long-range electromagnetic field
carried by the charged particle.  With a positive infrared cutoff it is a
well-defined Fock-space Weyl operator.  Removing the cutoff is the
representation-changing step described in the preceding section: the limiting
state is a charged hard datum together with a coherent soft electromagnetic
field in the corresponding asymptotic flux sector.  This is the leading
Faddeev--Kulish dressing datum.  Dressed-state and inclusive descriptions use
the same soft factor.  The inclusive construction sums over unresolved
radiation in probabilities, while the dressed construction builds part of the
soft cloud into the state representation.

\section{A Gauge-Invariant Infrared Regulator}

The cutoff \(\mu\) in the formulas above can be realized by deforming the
theory so that the photon has a small mass, and then removing the deformation.
If the same soft region is regulated dimensionally instead, the logarithms
\(\log(\mu/M)\) are replaced by infrared poles \(1/\epsilon_{\rm IR}\) plus
scheme-dependent finite constants; the coefficient multiplying the logarithm
or pole is the same eikonal coefficient \(A_{\beta\alpha}\), so the
real--virtual cancellation is identical coefficient by coefficient.
A convenient gauge-invariant deformation is an Abelian-Higgs regulator.
Introduce a complex scalar \(\varphi\), charged under the gauge field
\(A_\mu\), with
\[
  D_\mu\varphi=(\partial_\mu+\ii g_H A_\mu)\varphi
\]
and Lagrangian terms
\[
  \mathcal L_{\rm reg}
  =
  -\frac14F_{\mu\nu}F^{\mu\nu}
  -|D_\mu\varphi|^2
  -V(|\varphi|).
\]
Choose the potential so that its vacuum satisfies
\[
  |\vev{\varphi}|=v\ne0 .
\]
Expanding around this vacuum gives
\[
  -|D_\mu\varphi|^2
  \supset
  -g_H^2v^2 A_\mu A^\mu ,
\]
so the gauge boson acquires a mass proportional to \(g_Hv\).  Taking \(v\) to
zero after forming inclusive quantities gives a consistent version of the
infrared-regulated calculation.

Figure~\ref{fig:volume-ii-abelian-higgs-ir-regulator} records the Abelian-Higgs
infrared regulator used for this comparison: the gauge boson mass is generated
by a regulated Higgs expectation value and is taken to zero only after inclusive
quantities have been formed.

\begin{figure}[!ht]
  \centering
  \resizebox{0.9\textwidth}{!}{%
  \begin{tikzpicture}[>=Stealth,x=1cm,y=1cm,every node/.style={font=\small}]
    \begin{scope}
      \node[font=\normalsize] at (0,2.45) {regulator scalar};
      \draw[->,line width=0.85pt] (-3.0,0)--(3.1,0)
        node[below] {\(\operatorname{Re}\varphi\)};
      \draw[->,line width=0.85pt] (0,-0.28)--(0,2.1)
        node[left] {\(V\)};
      \draw[blue!70!black,very thick,domain=-2.35:2.35,samples=120]
        plot (\x,{0.23*(\x*\x-1.55)^2+0.08});
      \fill[green!50!black] (1.245,0.08) circle (2.4pt);
      \node[green!45!black,below] at (1.245,-0.12) {\(v\)};
      \node[align=center,blue!70!black] at (0,-0.75)
        {choose a vacuum\\\(|\vev{\varphi}|=v\)};
    \end{scope}

    \begin{scope}[shift={(6.3,0)}]
      \node[font=\normalsize] at (0,2.45) {massive photon regulator};
      \draw[line width=1.0pt] (-2.3,0.72)--(-0.6,0.72);
      \draw[purple!80!black,decorate,
        decoration={snake,amplitude=0.78mm,segment length=3.2mm},
        line width=0.95pt]
        (-0.6,0.72)--(1.25,0.72);
      \draw[line width=1.0pt] (1.25,0.72)--(2.4,0.72);
      \node[purple!80!black,below] at (0.34,0.47) {\(m_\gamma\propto g_Hv\)};
      \draw[->,line width=1.0pt] (0,0.3)--(0,-0.62);
      \node[align=center] at (0,-1.15)
        {\(-g_H^2v^2A_\mu A^\mu\)\\from \(|D_\mu\varphi|^2\)};
      \node[align=center,orange!80!black] at (0,1.78)
        {compute at \(v\ne0\),\\then take \(v\to0\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{A small photon mass may be introduced through a Higgs deformation
  while keeping gauge invariance manifest.}
  \label{fig:volume-ii-abelian-higgs-ir-regulator}
\end{figure}

The infrared analysis will be used later whenever massless gauge fields occur.
The next step in the logical development is the generating functional
formalism, which packages connected Green functions and one-particle
irreducible vertices in a form suitable for renormalization.
