\chapter{Jets, Infrared-Safe Observables, and Hadronization}
\label{ch:jets-ir-safe-observables-hadronization}
\ConventionsMixed

Jets enter quantum field theory as observables built from the energy and
momentum deposited by physical final states.  The final states in QCD are
hadronic states in the physical Hilbert space; quarks and gluons are
short-distance variables used inside a gauge-fixed or factorized calculation.
The mathematical question is therefore not how to assign reality to colored
partons, but when a measurement on physical final states admits a controlled
short-distance expansion in which partonic variables are legitimate internal
coordinates.

\section{Weighted Cross Sections and Measurement Functions}

Let \(\ket{i}\) be an incoming scattering state and let \(X\) range over a
complete family of outgoing physical states.  A detector observable is a
measurable function \(F(X)\) of the final state.  Its weighted inclusive cross
section has the form
\begin{equation}
  \sigma[F]
  =
  \frac{1}{\mathcal N_i}
  \sum_X
  (2\pi)^D\delta^{(D)}(P_i-P_X)
  |\langle X|T|i\rangle|^2\,F(X),
  \label{eq:jet-weighted-cross-section-hilbert}
\end{equation}
where \(\mathcal N_i\) is the incoming flux normalization and \(T\) is the
transition operator defined by \(S=\mathbf 1+\ii T\).  Equation
\eqref{eq:jet-weighted-cross-section-hilbert} is stated in terms of physical
states.  A perturbative partonic formula is a further expansion of the same
quantity, after an infrared regulator and an ultraviolet renormalization
scheme have been chosen:
\begin{equation}
  \sigma_{\rm part}[F]
  =
  \sum_{n\geq0}
  \int \dd\Phi_n\,
  |\mathcal M_n(p_1,\ldots,p_n)|^2\,
  F_n(p_1,\ldots,p_n).
  \label{eq:jet-partonic-measurement-function}
\end{equation}
Here \(\dd\Phi_n\) is the \(n\)-body Lorentz-invariant phase-space measure,
\(\mathcal M_n\) is a renormalized partonic matrix element, and \(F_n\) is the
partonic representative of the detector measurement.  The functions \(F_n\)
must satisfy compatibility conditions at the boundaries of phase space where a
parton becomes soft or two partons become collinear.

\begin{definition}[Infrared and collinear safety]
\label{def:irc-safe-measurement}
A family of measurement functions \(F=\{F_n\}_{n\geq0}\) is infrared and
collinear safe, abbreviated IRC safe, if the following limits hold on the
phase-space regions relevant to the observable:
\begin{align}
  \lim_{\lambda\downarrow0}
  F_{n+1}(p_1,\ldots,p_n,\lambda q)
  &=
  F_n(p_1,\ldots,p_n),
  \label{eq:irc-soft-safety}\\
  \lim_{\theta\downarrow0}
  F_{n+1}(p_1,\ldots,zp,(1-z)p,\ldots,p_n)
  &=
  F_n(p_1,\ldots,p,\ldots,p_n),
  \qquad 0<z<1.
  \label{eq:irc-collinear-safety}
\end{align}
The convergence is required with enough uniformity that the difference between
the two sides is integrable against the soft and collinear factorization
kernels of the theory at the perturbative order under consideration.
Equivalently, near a single unresolved boundary and away from other unresolved
boundaries, there must be constants \(C,\alpha>0\), depending on the compact
set of resolved momenta but not on the unresolved scale, such that
\begin{align}
  |F_{n+1}(p_1,\ldots,p_n,\lambda q)-F_n(p_1,\ldots,p_n)|
  &\le C\lambda^\alpha,\\
  |F_{n+1}(p_1,\ldots,zp,(1-z)p,\ldots,p_n)
    -F_n(p_1,\ldots,p,\ldots,p_n)|
  &\le C\theta^\alpha ,
\end{align}
where \(\theta\) is the splitting angle.  On corners where several emissions
are unresolved, the same estimate is required after replacing each unresolved
cluster by its reduced soft or collinear configuration.  These estimates make
the products with \(\dd\lambda/\lambda\) and
\(\dd\theta^2/\theta^2\) locally integrable after the real--virtual
subtraction.
\end{definition}

\begin{controlledapproximation}[Fixed-order cancellation for IRC-safe measurements]
\label{capprox:irc-safety-fixed-order-finiteness}
Consider an inclusive observable in a massless gauge theory for which the
renormalized real and virtual matrix elements admit the soft and collinear
boundary factorizations displayed below, with remainders locally integrable
after subtracting the reduced lower-multiplicity channel.  If the measurement
family \(F\) is IRC safe in the sense of
Definition~\ref{def:irc-safe-measurement}, then real-emission singularities
and virtual singularities cancel locally on the unresolved phase-space
boundaries after the KLN sum over degenerate final states.  With incoming
hadrons, the remaining initial-state collinear singularities are absorbed into
the renormalized parton distribution functions.
\end{controlledapproximation}

The derivation is a fixed-order perturbative one and uses the displayed
factorization formulae as input.  It is not a nonperturbative theorem about
arbitrary jet measurements.  In a soft region, with an extra gluon momentum
\(k=\lambda q\), the real matrix element factorizes as
\[
  |\mathcal M_{n+1}|^2\,\dd\Phi_{n+1}
  =
  |\mathcal M_n|^2\,\dd\Phi_n\,
  \left[
    \sum_{i,j}\mathcal S_{ij}(q)
  \right]
  \frac{\dd\lambda}{\lambda}\dd\Omega_q
  +R_{\rm soft}.
\]
The remainder \(R_{\rm soft}\) is locally integrable by the hypothesis of the
controlled approximation.
The eikonal factor \(\mathcal S_{ij}\) is the same factor that appears with
opposite sign in the virtual soft correction to the \(n\)-parton channel.
Equation \eqref{eq:irc-soft-safety} replaces the real measurement by
\(F_n\) on the singular part of the integral, so the real and virtual soft
integrals have the same weight and cancel.

In a final-state collinear region, two momenta approach
\[
  p_a=zp+r,\qquad p_b=(1-z)p-r,
  \qquad r_\perp^2\to0.
\]
The real matrix element and phase space factorize as
\[
  |\mathcal M_{n+1}|^2\,\dd\Phi_{n+1}
  =
  |\mathcal M_n(\ldots,p,\ldots)|^2\,\dd\Phi_n\,
  \frac{\alpha_s}{2\pi}
  P_{a b\leftarrow c}(z)
  \frac{\dd r_\perp^2}{r_\perp^2}\dd z
  +R_{\rm coll}.
\]
The remainder \(R_{\rm coll}\) is locally integrable by the hypothesis of the
controlled approximation.
Equation \eqref{eq:irc-collinear-safety} gives the same measurement weight to
the unresolved pair and to its parent momentum.  The singular integral then
cancels against the corresponding virtual collinear correction in the KLN
sum.  If the collinear parton is in the initial state of a hadron collision,
the same factorization kernel multiplies the operator renormalization of the
parton distribution function rather than canceling inside the hard partonic
subprocess.  The finite hard coefficient is obtained after this mass
factorization step.

Examples of IRC-safe observables include the total \(e^+e^-\) hadronic cross
section, thrust
\[
  T(X)
  =
  \max_{\widehat n}
  \frac{\sum_{h\in X}|\vec p_h\cdot\widehat n|}
       {\sum_{h\in X}|\vec p_h|},
\]
the \(C\)-parameter, jet masses defined with an IRC-safe algorithm, and the
energy correlators of
Chapter~\ref{ch:volume-ii-19}.  Observables that measure an identified
hadron, a charged-particle multiplicity, or a single unsummed collinear
fragment require additional nonperturbative input such as fragmentation
functions or detector-specific response functions.

\section{Sterman--Weinberg Jets}

The original two-jet observable of Sterman and Weinberg is a clean example of
the preceding definition.  In \(e^+e^-\) annihilation at center-of-mass energy
\(Q\), choose two parameters
\[
  0<\epsilon_{\rm out}<1,\qquad 0<\delta_{\rm cone}<\pi.
\]
An event is counted as a two-jet event if there exist two cones of half-angle
\(\delta_{\rm cone}\) such that the total energy outside the two cones is at
most \(\epsilon_{\rm out}Q\).  The corresponding measurement function is
\[
  F_{\rm SW}(X)
  =
  \begin{cases}
  1,& \text{if }X\text{ satisfies the two-cone energy condition},\\
  0,& \text{otherwise}.
  \end{cases}
\]

Soft safety follows because adding a particle of energy \(\lambda Q\) changes
the out-of-cone energy by at most \(\lambda Q\).  Away from the boundary
where the out-of-cone energy is exactly \(\epsilon_{\rm out}Q\), the
classification is unchanged for sufficiently small \(\lambda\).  The boundary
has zero measure in the phase-space integral, and the unresolved soft
singularity is integrated with the same measurement weight as the event
without the soft particle.

Collinear safety follows because a sufficiently collinear splitting of a
momentum inside one of the cones remains inside the same cone, while a
splitting outside the cones preserves the total out-of-cone energy in the
collinear limit.  The singular collinear phase-space integral therefore sees
the parent momentum measurement.

For very small \(\epsilon_{\rm out}\) and \(\delta_{\rm cone}\), the fixed-order
series contains logarithms such as
\[
  \alpha_s^m
  \log^r\epsilon_{\rm out}\,
  \log^s\delta_{\rm cone},
  \qquad r+s\leq 2m.
\]
The two-jet cross section is then still an IRC-safe observable, but the useful
short-distance expansion is a resummed expansion rather than a fixed-order
polynomial in \(\alpha_s(Q)\).

\section{Sequential Recombination Jet Algorithms}

Modern collider jet definitions are usually algorithmic.  For a hadron
collider, let \(p_{Ti}\) be transverse momentum, \(y_i\) rapidity, and
\(\phi_i\) azimuth.  Define
\[
  \Delta R_{ij}^2=(y_i-y_j)^2+(\phi_i-\phi_j)^2.
\]
The generalized \(k_T\) family with radius \(R\) and exponent \(p\) uses
\begin{equation}
  d_{ij}
  =
  \min(p_{Ti}^{2p},p_{Tj}^{2p})
  \frac{\Delta R_{ij}^2}{R^2},
  \qquad
  d_{iB}=p_{Ti}^{2p}.
  \label{eq:generalized-kt-distance}
\end{equation}
At each step the smallest number among all \(d_{ij}\) and \(d_{iB}\) is
selected.  If it is \(d_{ij}\), the two objects are recombined, usually by
adding their four-momenta.  If it is \(d_{iB}\), object \(i\) is declared a
jet and removed from the list.  The cases
\[
  p=1,\qquad p=0,\qquad p=-1
\]
give the \(k_T\), Cambridge--Aachen, and anti-\(k_T\) algorithms,
respectively.

Figure~\ref{fig:jets-measurement-irc-flow} shows the geometry behind
\eqref{eq:generalized-kt-distance}.  The same rapidity--azimuth event can
produce a circular hard catchment region for anti-\(k_T\), while the \(k_T\)
metric permits soft constituents to precluster and therefore gives a
catchment boundary controlled by soft radiation.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,>=Stealth,every node/.style={font=\scriptsize}]
    \begin{scope}
      \draw[->] (-2.2,-1.65)--(2.25,-1.65) node[right] {\(y\)};
      \draw[->] (-2.2,-1.65)--(-2.2,1.75) node[above] {\(\phi\)};
      \node[font=\small] at (0,2.05) {anti-\(k_T\), \(p=-1\)};

      \draw[blue!65!black,thick] (0,0) circle (1.25);
      \node[blue!65!black] at (1.33,0.12) {\(\Delta R_{h s}=R\)};
      \fill[blue!70!black] (0,0) circle (4pt);
      \node[below right] at (0.05,-0.05) {\(h,\ p_{Th}\)};
      \fill[orange!80!black] (0.75,0.55) circle (2pt);
      \fill[orange!80!black] (-0.62,0.35) circle (1.7pt);
      \fill[orange!80!black] (0.24,-0.91) circle (1.6pt);
      \fill[orange!80!black] (1.52,-0.22) circle (1.7pt);
      \fill[orange!80!black] (-1.48,-0.43) circle (1.5pt);
      \fill[black!65] (1.88,0.82) circle (2.8pt);
      \node[above right] at (1.88,0.82) {other hard};

      \draw[densely dashed,orange!80!black] (0,0)--(0.75,0.55);
      \node[align=center,text width=4.2cm] at (0,-2.25)
      {\(d_{h s}\simeq p_{Th}^{-2}\Delta R_{h s}^{2}/R^2,\quad d_{hB}=p_{Th}^{-2}\).\\
       If \(\Delta R_{h s}<R\), then \(d_{h s}<d_{hB}\).};
    \end{scope}

    \begin{scope}[shift={(6.0,0)}]
      \draw[->] (-2.2,-1.65)--(2.25,-1.65) node[right] {\(y\)};
      \draw[->] (-2.2,-1.65)--(-2.2,1.75) node[above] {\(\phi\)};
      \node[font=\small] at (0,2.05) {\(k_T\), \(p=1\)};

      \draw[purple!70!black,thick,dashed]
        plot[smooth cycle,tension=0.75] coordinates
        {(-1.05,-0.72) (-0.35,-1.05) (0.62,-0.84) (1.05,-0.18)
         (0.86,0.72) (0.16,1.03) (-0.68,0.64)};
      \fill[blue!70!black] (0,0) circle (4pt);
      \node[below right] at (0.05,-0.05) {\(h,\ p_{Th}\)};
      \fill[orange!80!black] (0.75,0.55) circle (2pt);
      \fill[orange!80!black] (-0.62,0.35) circle (1.7pt);
      \fill[orange!80!black] (0.24,-0.91) circle (1.6pt);
      \fill[orange!80!black] (1.52,-0.22) circle (1.7pt);
      \fill[orange!80!black] (-1.48,-0.43) circle (1.5pt);
      \fill[black!65] (1.88,0.82) circle (2.8pt);
      \node[above right] at (1.88,0.82) {other hard};

      \draw[orange!80!black,densely dashed] (-0.62,0.35)--(-1.48,-0.43)--(0.24,-0.91);
      \draw[orange!80!black,densely dashed] (0.75,0.55)--(1.52,-0.22);
      \node[align=center,text width=4.2cm] at (0,-2.25)
      {\(d_{ss'}=\min(p_{Ts}^2,p_{Ts'}^2)\Delta R_{ss'}^{2}/R^2\).\\
       Soft--soft distances can be smallest.};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{Generalized \(k_T\) clustering in the rapidity--azimuth plane.  For
  anti-\(k_T\), a soft constituent inside \(\Delta R<R\) clusters with the
  hard particle before the hard particle reaches the beam distance, producing
  an approximately circular hard catchment region.  For \(k_T\), soft--soft
  distances may be smallest, so soft preclusters can control the eventual jet
  boundary even for the same event.}
  \label{fig:jets-measurement-irc-flow}
\end{figure}

For \(p=1,0,-1\), the generalized \(k_T\) clustering algorithm with
four-momentum recombination defines IRC-safe jet momenta and IRC-safe
cross sections after imposing a finite jet transverse-momentum threshold.
For a collinear splitting \(p\to zp,(1-z)p\), the angular distance satisfies
\(\Delta R_{ab}\to0\), so \(d_{ab}\to0\).  The two collinear objects are
therefore recombined before they can affect finite-angle clustering, and
four-momentum recombination restores the parent momentum.

For a soft momentum \(q\), the effect of recombination with a hard object
changes the hard jet momentum by \(O(q)\).  If the soft object is clustered
among other soft objects or declared as a separate jet, a finite jet
transverse-momentum threshold removes it in the soft limit.  In the anti-\(k_T\)
case, a soft object within radius \(R\) of a hard jet has pair distance
controlled by the hard transverse momentum, so it is pulled into the hard jet
without moving the hard boundary at leading power.  The \(k_T\) and
Cambridge--Aachen cases may cluster soft radiation differently, but the
resulting change of any finite-energy jet momentum vanishes with the soft
energy.  Thus the measurement functions associated with the algorithm obey
\eqref{eq:irc-soft-safety} and \eqref{eq:irc-collinear-safety}.

The anti-\(k_T\) algorithm is widely used because hard jets acquire stable,
approximately circular catchment areas in the rapidity--azimuth plane while
retaining IRC safety.  Cone algorithms require more care.  A seedless
infrared-safe cone algorithm can be defined by searching over all stable
cones.  A seed-based cone algorithm with a transverse-momentum seed threshold
can fail soft safety: adding an arbitrarily soft seed may create a stable cone
that was absent before, changing the hard jets by a finite amount.

\section{Factorization and Resummation}

IRC safety makes fixed-order perturbation theory finite, but it does not make
every region of the observable distribution well approximated by a few fixed
orders.  When an event shape \(e(X)\) is small, the final state is constrained
to contain narrow energetic jets plus soft radiation.  A factorized
description separates the scales, but the separation is not a statement that
the exact QCD Hilbert space has literally split into independent colored
subspaces.  It is a statement about a controlled expansion of a
gauge-invariant weighted cross section in which the short-distance coefficient
and the long-distance matrix elements are separately defined after a regulator,
an overlap-subtraction prescription, and a renormalization scheme are chosen.

For a two-jet event shape in \(e^+e^-\), a leading singular factorized form has
the convolution form
\begin{equation}
  \frac{\dd\sigma}{\dd e}
  =
  H(Q,\mu)
  \int \dd s_1\,\dd s_2\,\dd k\,
  J(s_1,\mu)J(s_2,\mu)S(k,\mu)\,
  \delta\!\left(e-\frac{s_1+s_2}{Q^2}-\frac{k}{Q}\right)
  +
  O(e^0).
  \label{eq:jet-event-shape-factorization}
\end{equation}
The displayed \(O(e^0)\) denotes terms that are nonsingular as distributions at
the two-jet endpoint; it is not an error estimate until the observable, test
function, and endpoint scaling region have been specified.  The hard function
\(H\) is the squared Wilson coefficient of the QCD current matched onto an
effective current.  The jet functions \(J\) are regulated collinear matrix
elements of gauge-invariant collinear fields.  The soft function \(S\) is a
regulated vacuum matrix element of soft Wilson lines with the soft measurement
operator inserted.  Their renormalization-group equations resum logarithms of
the separated scales.

Soft-collinear effective theory gives a systematic operator language for
factorizations of the form
\eqref{eq:jet-event-shape-factorization}.  Its assumptions are visible in the
operator definitions: mode scaling, Wilson-line decoupling at leading power,
measurement factorization, and a power expansion in the small event-shape
parameter.  The same logic applies to hadron-collider observables, with the
additional parton distribution functions needed for incoming hadrons and with
factorization-breaking channels treated as separate hypotheses.

\subsection{Mode Expansion and Logarithmic Accuracy}

For a two-jet observable with small dimensionless event shape \(e\), choose
lightlike vectors \(n,\bar n\) with \(n\cdot\bar n=2\) and write momenta as
\[
  p^\mu=\frac{n\cdot p}{2}\bar n^\mu
       +\frac{\bar n\cdot p}{2}n^\mu+p_\perp^\mu .
\]
The SCET expansion parameter \(\lambda\) is fixed by the measured region.  A
mode is not an exact particle species; it is a coordinate chart on momentum
regions, implemented in a regulator by smooth cutoffs or by an equivalent
subtraction prescription.  For thrust-like two-jet kinematics,
\(e\sim\lambda^2\), and the momentum scalings are
\[
  p_n^\mu\sim Q(\lambda^2,1,\lambda),\qquad
  p_{\bar n}^\mu\sim Q(1,\lambda^2,\lambda),\qquad
  p_s^\mu\sim Q(\lambda^2,\lambda^2,\lambda^2),
\]
where the three entries denote
\((n\cdot p,\bar n\cdot p,p_\perp)\).  A leading-power factorization claim
is a statement that, for \(e\ll1\) and away from additional measured
hierarchies,
\[
  \frac{\dd\sigma}{\dd e}
  =
  H\otimes J_n\otimes J_{\bar n}\otimes S
  +
  O(\lambda)
  \quad\text{or}\quad
  O(e^{1/2}),
\]
with the precise power depending on the measurement and recoil convention.
The symbol \(\otimes\) denotes the convolution variables in the displayed
factorization formula, not an abstract tensor product.  The hard function is a
Wilson coefficient of the QCD current matched onto SCET operators; the jet
functions are matrix elements of collinear fields with collinear Wilson
lines; the soft function is a vacuum matrix element of soft Wilson lines with
the measurement operator inserted.

\begin{definition}[SCET operator specification for a two-jet endpoint]
\label{def:scet-two-jet-operator-datum}
A two-jet SCET operator specification for a color-singlet source consists of:
\begin{enumerate}
  \item two lightlike directions \(n,\bar n\) with \(n\cdot\bar n=2\), a small
  parameter \(\lambda\), and momentum regions with the scaling displayed
  above;
  \item gauge-covariant collinear fields in the \(n\) and \(\bar n\) sectors,
  with collinear Wilson lines \(W_n,W_{\bar n}\) along the conjugate lightlike
  directions, and gauge-invariant building blocks such as
  \[
    \chi_n:=W_n^\dagger\xi_n,\qquad
    \mathcal B_{n\perp}^{\mu}:=W_n^\dagger\,\ii D_{n\perp}^{\mu}W_n
  \]
  where \(D_{n\perp}=\partial_{n\perp}-\ii A_{n\perp}\) uses the
  coupling-absorbed Yang--Mills connection of
  Chapter~\ref{ch:vol2-classical-yang-mills-matter}.  Thus the normalization
  of \(\mathcal B_{n\perp}\) is fixed by the same gauge-field coordinate that
  appears in
  \(-\frac1{4g_{\rm YM}^2}\operatorname{tr}(F_{\mu\nu}F^{\mu\nu})\);
  \item soft Wilson lines \(Y_n,Y_{\bar n}\) in the representations carried by
  the energetic charges, with their future- or past-pointing orientation fixed
  by whether the corresponding charge is outgoing or incoming;
  \item a matched current, for example
  \[
    J_{\rm QCD}(0)
    =
    C(Q,\mu)\,
    \bar\chi_{\bar n}(0)\Gamma\chi_n(0)
    +O(\lambda),
  \]
  before the soft Wilson-line decoupling is applied;
  \item renormalized jet and soft matrix elements, with the two-jet soft
  matrix element having the Wilson-line form
  \[
    S(k,\mu)
    =
    \frac1{N_c}
    \operatorname{tr}
    \bra{\vac}
    \overline{\operatorname T}\!\left(Y_{\bar n}^\dagger Y_n\right)
    \delta\!\left(k-\widehat k_s\right)
    \operatorname T\!\left(Y_n^\dagger Y_{\bar n}\right)
    \ket{\vac}_{\mu},
  \]
  where \(\widehat k_s\) is the soft measurement operator appropriate to the
  event shape;
  \item an overlap-subtraction prescription, often called a zero-bin
  subtraction, preventing the same momentum region from being counted both as
  soft and as the soft limit of a collinear mode;
  \item rapidity-regulator and rapidity-renormalization data whenever the
  Wilson-line geometry produces rapidity divergences not removed by ordinary
  ultraviolet renormalization;
  \item a statement about Glauber exchange: cancellation, absorption into an
  additional operator, or failure of the proposed factorization.
\end{enumerate}
All entries are part of the definition of the factorized approximation.  The
symbols \(H,J,S\) in \eqref{eq:jet-event-shape-factorization} have no invariant
meaning until these choices, including their renormalization scheme, have been
specified.
\end{definition}

\paragraph{Overlap subtraction as finite inclusion--exclusion.}
The zero-bin entry in Definition~\ref{def:scet-two-jet-operator-datum} is a
mathematical subtraction, not a phrase meaning that an integral may be ignored.
At finite regulator, let \(\Lambda\) be a finite set of momentum cells and let
\(\Lambda_n,\Lambda_s\subset\Lambda\) denote the cells represented by the
leading \(n\)-collinear and soft approximations.  Their common soft-collinear
corner is
\[
  \Lambda_0=\Lambda_n\cap\Lambda_s .
\]
Let \(f_\ell\) be the value of the measurement test function on cell
\(\ell\), and let \(I_n(\ell)\), \(I_s(\ell)\), and \(I_0(\ell)\) be the
regulated leading collinear, leading soft, and common overlap densities on
their domains.  The finite overlap-subtracted singular coordinate is
\begin{equation}
  I_{\rm scet}[f]
  =
  \sum_{\ell\in\Lambda_n} I_n(\ell)f_\ell
  +
  \sum_{\ell\in\Lambda_s} I_s(\ell)f_\ell
  -
  \sum_{\ell\in\Lambda_0} I_0(\ell)f_\ell .
  \label{eq:finite-scet-zero-bin-inclusion-exclusion}
\end{equation}
If \(I_n=I_s=I_0\) on \(\Lambda_0\), the overlap contribution is counted once,
not twice.  This is the finite inclusion--exclusion content of the zero-bin
subtraction.  In dimensional regularization some overlap integrals are
scaleless and vanish after analytic continuation; that is a feature of that
regulator and subtraction scheme, not a regulator-independent proof that the
overlap was absent.

A change of overlap convention is a finite scheme change only when it is
paired across the collinear and soft coordinates.  If
\(\delta I_0\) is moved from the collinear subtraction into the soft
coordinate, the two displayed sums in
\eqref{eq:finite-scet-zero-bin-inclusion-exclusion} change separately, but
their sum changes by
\[
  -\sum_{\ell\in\Lambda_0}\delta I_0(\ell)f_\ell
  +
  \sum_{\ell\in\Lambda_0}\delta I_0(\ell)f_\ell
  =
  0 .
\]
This is the finite algebraic reason that overlap subtraction data belong in
the definition of the factorized approximation.  The continuum theorem still
requires estimates showing that the regulated partition, the overlap
approximants, and the measurement remainder converge in the topology in which
the cross section is tested.

\paragraph{A regulated one-emission endpoint observable.}
The next calculation is a positive fixed-order case, not an all-orders SCET
theorem.  Take the color-singlet two-jet current and its one-gluon
soft-collinear singular graph.  In light-cone variables
\[
  u=\frac{n\cdot k}{Q},\qquad
  v=\frac{\bar n\cdot k}{Q},
\]
tilt the two Wilson-line velocities off the light cone by
\[
  \beta_n=n+\delta\bar n,\qquad
  \beta_{\bar n}=\bar n+\delta n,\qquad
  0<\delta<1 .
\]
For a massless cut gluon, the on-shell phase space after the transverse
delta function is a constant multiple of \(\dd u\,\dd v\).  The eikonal graph
has
\[
  \frac{\beta_n\cdot\beta_{\bar n}}
       {(\beta_n\cdot k)(\beta_{\bar n}\cdot k)}
  =
  \frac{2(1+\delta^2)}
       {Q^2(u+\delta v)(v+\delta u)} .
\]
Thus the off-light-cone regulator mixes the two light-cone components in the
denominators.  It is not the replacement \(u\mapsto u+\lambda\) and
\(v\mapsto v+\lambda\).  To keep the real-emission functional finite before
combining it with virtual terms, impose the endpoint slicing
\[
  \mathcal D_{\lambda,\rho}
  =
  \{(u,v):0\le u,v\le\rho,\ u+v\ge\lambda\},
  \qquad
  0<\lambda<\rho<\frac12 .
\]
The tilt \(\delta\) regulates rapidity; the lower slice \(\lambda\) regulates
the unresolved soft point of this real graph.  Smear the measured event-shape
variable \(e=u+v\) by \(\varphi\in C^2([0,2\rho])\).  Up to the common color,
coupling, and phase-space constant, the regulated one-emission functional is
\begin{equation}
  \Sigma_{\delta,\lambda,\rho}^{(1)}[\varphi]
  =
  \int_{\mathcal D_{\lambda,\rho}}
  K_\delta(u,v)\,\varphi(u+v)
  \,\dd u\,\dd v .
  \label{eq:regulated-scet-one-emission-functional}
\end{equation}
Here
\[
  K_\delta(u,v)
  =
  \frac{1+\delta^2}{(u+\delta v)(v+\delta u)} .
\]
This is the finite-regulator version of the specified soft Wilson-line cut
graph paired with a specified measurement; the graph, regulator, endpoint
window, and measurement all belong to the datum.  The domain
\(\mathcal D_{\lambda,\rho}\) is a soft endpoint domain: both light-cone
components are small.  In the SCET convention above,
\[
  p_n\sim Q(\lambda_{\rm SCET}^2,1,\lambda_{\rm SCET}),
  \qquad
  p_{\bar n}\sim Q(1,\lambda_{\rm SCET}^2,\lambda_{\rm SCET}),
  \qquad
  p_s\sim Q(\lambda_{\rm SCET}^2,\lambda_{\rm SCET}^2,
  \lambda_{\rm SCET}^2),
\]
so a genuine energetic \(n\)-collinear region has
\(\bar n\cdot k/Q=O(1)\), a genuine energetic \(\bar n\)-collinear region has
\(n\cdot k/Q=O(1)\), and neither is contained in
\(\mathcal D_{\lambda,\rho}\).  For the simple measurement \(e=u+v\), the
leading energetic-collinear term would be \(v\) in the \(n\)-collinear scaling
and \(u\) in the \(\bar n\)-collinear scaling; the endpoint Taylor projections
below do not establish those terms as jet matrix elements.  The symbol
\(\lambda\) in
\(\mathcal D_{\lambda,\rho}\) is only the lower endpoint slice of this finite
real graph, not the SCET expansion parameter.

Let
\[
  S_u[\varphi]
  :=
  \int_{\mathcal D_{\lambda,\rho}}
  K_\delta(u,v)\,\varphi(u)\,\dd u\,\dd v,
  \qquad
  S_v[\varphi]
  :=
  \int_{\mathcal D_{\lambda,\rho}}
  K_\delta(u,v)\,\varphi(v)\,\dd u\,\dd v,
\]
and
\[
  S_0[\varphi]
  :=
  \varphi(0)
  \int_{\mathcal D_{\lambda,\rho}}K_\delta(u,v)\,\dd u\,\dd v .
\]
These three terms are obtained from the same regulated soft graph by
Taylor-projection of the measurement on the endpoint square.  They are useful
finite coordinates on the soft endpoint singularity, but they are not the
\(n\)-jet, \(\bar n\)-jet, and method-of-regions zero-bin matrix elements of a
full SCET factorization theorem.  In a genuine region derivation the
\(n\)-collinear, \(\bar n\)-collinear, hard, and soft graphs live on domains
with the scalings just displayed, and the overlap subtraction is derived by
expanding those distinct approximants on their common limit.  The
constant-subtracted soft endpoint coordinate is
\begin{equation}
  \Sigma_{\delta,\lambda,\rho}^{\rm soft\,proj}[\varphi]
  =
  S_u[\varphi]+S_v[\varphi]-S_0[\varphi].
  \label{eq:regulated-scet-one-emission-subtracted}
\end{equation}
Equivalently, for any auxiliary allocation parameter \(a\),
\[
  \bigl(S_u-aS_0\bigr)
  +
  \bigl(S_v-(1-a)S_0\bigr)
  =
  S_u+S_v-S_0,
\]
so this algebraic soft projection is independent of how the constant endpoint
term is split between the two Taylor projections.  Omitting the subtraction
changes the answer by the nonzero constant endpoint term \(S_0\).  This
identity is the finite shadow of zero-bin inclusion--exclusion, not a
stand-alone derivation of the collinear operators.

The exact regulated graph decomposes as
\begin{equation}
  \Sigma_{\delta,\lambda,\rho}^{(1)}[\varphi]
  =
  \Sigma_{\delta,\lambda,\rho}^{\rm soft\,proj}[\varphi]
  +
  R_{\delta,\lambda,\rho}^{\rm mix}[\varphi],
  \label{eq:regulated-scet-one-emission-decomposition}
\end{equation}
where the mixed endpoint remainder is
\begin{equation}
  R_{\delta,\lambda,\rho}^{\rm mix}[\varphi]
  =
  \int_{\mathcal D_{\lambda,\rho}}
  K_\delta(u,v)
  \bigl[\varphi(u+v)-\varphi(u)-\varphi(v)+\varphi(0)\bigr]
  \,\dd u\,\dd v .
  \label{eq:regulated-scet-one-emission-remainder}
\end{equation}
Indeed
\[
  \varphi(u+v)
  =
  \varphi(u)+\varphi(v)-\varphi(0)
  +
  \bigl[\varphi(u+v)-\varphi(u)-\varphi(v)+\varphi(0)\bigr],
\]
and the first three terms integrate to
\eqref{eq:regulated-scet-one-emission-subtracted}.  If
\(M_\varphi=\sup_{0\le t\le2\rho}|\varphi''(t)|\), then
\[
  \varphi(u+v)-\varphi(u)-\varphi(v)+\varphi(0)
  =
  \int_0^u\!\!\int_0^v \varphi''(s+t)\,\dd t\,\dd s,
\]
so
\begin{equation}
  |R_{\delta,\lambda,\rho}^{\rm mix}[\varphi]|
  \le
  M_\varphi
  \int_{\mathcal D_{\lambda,\rho}}K_\delta(u,v)uv\,\dd u\,\dd v
  \le
  M_\varphi(1+\delta^2)\rho^2 .
  \label{eq:regulated-scet-one-emission-remainder-bound}
\end{equation}
The last inequality uses
\((u+\delta v)(v+\delta u)\ge uv\) and
\({\rm area}(\mathcal D_{\lambda,\rho})\le\rho^2\).  This displays the
soft Wilson-line endpoint mechanism actually proved by the simple graph: the
constant endpoint projection multiplies \(\varphi(0)\), the two first
projections test the \(u\)- and \(v\)-dependence of the same soft graph, and
the mixed finite-difference remainder is controlled in the same
smeared-observable topology.  What has not been proved here is equally
important: the graph does not contain the energetic collinear domains and does
not produce the hard coefficient.  The all-orders continuum QCD factorization
theorem still requires the full operator construction, separate hard and
collinear region expansions, regulator removal, rapidity renormalization,
Glauber analysis, and a power remainder for the physical cross section.

\paragraph{Leading soft Wilson-line decoupling identity.}
At finite regulator and at leading power in the \(n\)-collinear sector, the
soft connection appearing through \(n\cdot D_s=n\cdot\partial-\ii n\cdot A_s\)
can be removed from the collinear kinetic operator by multiplication by a soft
Wilson line.  More precisely, if \(Y_n\) satisfies
\[
  n\cdot D_s\,Y_n=0,
\]
and the original collinear field is written \(\xi_n=Y_n\xi_n^{(0)}\), then
\[
  \ii n\cdot D_s\,\xi_n
  =
  Y_n\,\ii n\cdot\partial\,\xi_n^{(0)} .
\]
Thus the leading soft interactions are transferred from the collinear
Lagrangian to Wilson lines in the external operators and measurement matrix
elements.

Using the Leibniz rule,
\[
  n\cdot D_s(Y_n\xi_n^{(0)})
  =
  (n\cdot D_sY_n)\xi_n^{(0)}
  +Y_n\,n\cdot\partial\,\xi_n^{(0)} .
\]
The first term vanishes by the defining differential equation for \(Y_n\).
Multiplying by \(\ii\) gives the displayed identity.  In a finite-dimensional
regulated path integral, the change of variables is multiplication by an
invertible color matrix at each regulated spacetime point.  With a
gauge-covariant regulator its Berezin or bosonic Jacobian is a local
gauge-invariant factor; in the factorized construction that factor is included
in the same renormalized soft and collinear operator specification rather than treated
as an extra physical assumption.  The identity itself is algebraic and does
not assert that subleading soft interactions or Glauber exchanges decouple.

\paragraph{Finite Glauber unitarity diagnostic.}
Glauber exchange is a factorization boundary because its momentum scaling is
not removed by the leading soft Wilson-line identity above.  At finite
regulator one useful diagnostic is purely operator-theoretic.  Let
\(\mathcal H_{\rm spec}\) be the finite Hilbert space carrying the spectator
and exchanged color coordinates that are not resolved by a proposed
inclusive measurement, let \(\rho_{\rm spec}\) be the corresponding reduced
density matrix, and let \(U_G\) be the finite unitary generated by the
regulated Glauber operator on this space.  If the measurement operator is
inclusive over these coordinates, so that it is represented by
\(M_{\rm incl}=\mathbf 1_{\rm spec}\) on this factor, then
\begin{equation}
  \operatorname{Tr}_{\rm spec}
  \!\left(M_{\rm incl}U_G\rho_{\rm spec}U_G^\dagger\right)
  =
  \operatorname{Tr}_{\rm spec}\rho_{\rm spec}.
  \label{eq:glauber-inclusive-unitarity-diagnostic}
\end{equation}
More generally, the same finite cancellation holds whenever the measured
operator \(M\) commutes with \(U_G\):
\begin{equation}
  \operatorname{Tr}_{\rm spec}\!\left(MU_G\rho_{\rm spec}U_G^\dagger\right)
  =
  \operatorname{Tr}_{\rm spec}\!\left(U_G^\dagger M U_G\rho_{\rm spec}\right)
  =
  \operatorname{Tr}_{\rm spec}(M\rho_{\rm spec}).
  \label{eq:glauber-commuting-measurement-diagnostic}
\end{equation}
Thus an inclusive color-singlet measurement may be insensitive to a unitary
Glauber rotation even though the rotation is present at amplitude level.  The
same algebra also shows where a factorization assertion can fail.  If the
measurement projects onto a spectator or color-flow coordinate that does not
commute with \(U_G\), then \(U_G^\dagger M U_G\neq M\), and the trace need not
equal the value without the Glauber exchange.  In that case the exchanged
Glauber coordinate must either be included as an additional operator in the
factorization structure or the proposed factorization statement is false for that
measurement.

There is also a quantitative finite version, useful because real
factorization proofs rarely give exact commutation at an intermediate
regulator.  For any bounded finite measurement \(M\), define the Glauber
remainder
\begin{equation}
  \Delta_G(M,\rho_{\rm spec})
  =
  \operatorname{Tr}_{\rm spec}
  \!\left(MU_G\rho_{\rm spec}U_G^\dagger\right)
  -
  \operatorname{Tr}_{\rm spec}(M\rho_{\rm spec}).
  \label{eq:glauber-finite-remainder-definition}
\end{equation}
Cyclicity gives the exact identity
\begin{equation}
  \Delta_G(M,\rho_{\rm spec})
  =
  \operatorname{Tr}_{\rm spec}
  \!\left((U_G^\dagger M U_G-M)\rho_{\rm spec}\right).
  \label{eq:glauber-finite-remainder-commutator}
\end{equation}
Consequently, with \(\|\cdot\|\) the operator norm and
\(\|\cdot\|_1\) the trace norm,
\begin{equation}
  |\Delta_G(M,\rho_{\rm spec})|
  \le
  \|U_G^\dagger M U_G-M\|\,\|\rho_{\rm spec}\|_1 .
  \label{eq:glauber-finite-remainder-bound}
\end{equation}
If \(U_G=\exp(\ii K_G)\), the Duhamel formula refines the first factor to
\[
  U_G^\dagger M U_G-M
  =
  \ii\int_0^1
  e^{-\ii sK_G}[M,K_G]e^{\ii sK_G}\,\dd s ,
\]
and hence
\[
  \|U_G^\dagger M U_G-M\|
  \le
  \|[M,K_G]\| .
\]
For a concrete breaking example take
\[
  U_G=
  \begin{pmatrix}
    \cos\theta&\sin\theta\\
    -\sin\theta&\cos\theta
  \end{pmatrix},
  \qquad
  M=
  \begin{pmatrix}
    1&0\\0&0
  \end{pmatrix},
  \qquad
  \rho_{\rm spec}=
  \begin{pmatrix}
    r_1&0\\0&r_2
  \end{pmatrix}.
\]
Then
\[
  \Delta_G(M,\rho_{\rm spec})
  =
  (r_2-r_1)\sin^2\theta .
\]
Unless \(r_1=r_2\) or \(\theta\in\pi\mathbb Z\), the measurement detects the
Glauber rotation.  The proposed factorized expression must then carry this
Glauber operator or restrict the measurement; allocating a residual symbol
does not by itself prove that the residual is small.
The finite diagnostic therefore has an honest error coordinate: exact
factorization is recovered when the measured algebra is invariant under the
Glauber unitary, and approximate factorization requires a small commutator in
the measurement norm relevant to the smeared cross section.  This is the
operator-theoretic content behind phrases such as "Glauber cancellation";
the continuum theorem must still supply the limiting \(K_G\), the topology,
and the bound.

Equations
\eqref{eq:glauber-inclusive-unitarity-diagnostic}--\eqref{eq:glauber-commuting-measurement-diagnostic}
are only the finite unitarity mechanism.  A continuum QCD factorization proof
must also identify the regulated Glauber region, show how the eikonal
approximations and contour deformations isolate it, prove the required
inclusive sum or commutation property for the actual measurement, and bound
the remainder in the topology of the smeared cross section.  Without these
extra analytic steps, saying that Glauber exchanges ``cancel'' is merely a
claim about a missing part of the factorization structure.

\begin{controlledapproximation}[Spectator-model obstruction to generalized TMD factorization]
\label{ca:scet-spectator-tmd-color-entanglement}
The finite diagnostic above becomes a real factorization-breaking calculation
in the nonabelian spectator model of Rogers and Mulders
\cite{RogersMulders2010NoGeneralizedTMD}.  The measured process is
\[
  H_1(P_1,s_{1\perp})+H_2(P_2,s_{2\perp})
  \longrightarrow
  h_3(k_3)+h_4(k_4)+X ,
  \qquad
  q_\perp:=k_{3\perp}+k_{4\perp},\quad |q_\perp|\ll Q ,
\]
with the cross section kept differential in the recoil \(q_\perp\) and in
the double transverse-spin projection.  The active partons are collinear to
the incoming hadrons, \(k_1^+\sim P_1^+\) and \(k_2^-\sim P_2^-\), while the
exchanged gluons that generate the obstruction are in the Glauber region
\(|l_i^+|,|l_i^-|\ll |l_{i\perp}|\).  A color-neutral hard boson is used only
to remove irrelevant hard color routings; the nonabelian color is carried by
the initial and final colored lines.

The generalized TMD ansatz ruled out by the calculation is the statement that
all leading recoil dependence can be written, even with process-dependent
Wilson lines, as a product of separate one-hadron and one-fragmentation
matrix elements:
\begin{equation}
  \dd\sigma_{\rm gen}
  \stackrel{?}{=}
  \sum_c
  H_c
  \otimes
  \Phi_{1}^{[W_1^c]}
  \otimes
  \Phi_{2}^{[W_2^c]}
  \otimes
  \Delta_{3}^{[W_3^c]}
  \otimes
  \Delta_{4}^{[W_4^c]}
  +
  O(q_\perp/Q),
  \label{eq:scet-generalized-tmd-ansatz}
\end{equation}
where the Wilson-line geometry may depend on the hard color routing \(c\), but
there is no joint \(H_1H_2\) soft or Glauber matrix element.  This ansatz is
stronger than ordinary collinear factorization and is not the Drell--Yan
color-singlet TMD statement.

At one extra gluon, the would-be anomalous eikonal term on a separate
color-traced loop is
\begin{equation}
  \mathcal E_1^a(l)
  =
  -2\ii\pi g\,n_1^\mu\,\delta(l^+)\,
  \operatorname{tr}(t^a)
  =
  0 .
  \label{eq:scet-spectator-single-gluon-trace-zero}
\end{equation}
Thus a separate TMD factor may not be given a nonzero order-\(g\) anomalous
term without contradicting the graph sum.  At two gluons emitted from the
same hadron, \(\operatorname{tr}(t^at^b)\) is nonzero and the anomaly can be
represented by a color-traced Wilson loop inside that hadron's TMD.  This is
the largest repair allowed by the separate-factor ansatz.

Now attach one Glauber gluon to the \(H_1\)-collinear spectator system and
one to the \(H_2\)-collinear spectator system.  The color factor in the graph
sum is
\begin{equation}
  C_{\rm ent}
  =
  \operatorname{tr}(t^at^b)\operatorname{tr}(t^bt^a)
  =
  T_F^2(N_c^2-1),
  \label{eq:scet-spectator-color-entanglement-factor}
\end{equation}
which is nonzero for \(SU(N_c)\).  The cut structure also does not cancel it.
For the graphs with both extra gluons on the same side of the final-state
cut, the eikonal denominators reduce to
\begin{align}
  &\left( \frac{1}{-l_1^+ + \ii0} \right)
   \left( \frac{1}{-l_2^- + \ii0} \right)
  +\left( \frac{1}{ l_1^+ + \ii0} \right)
   \left( \frac{1}{-l_2^- + \ii0} \right)
  \notag\\
  &\quad
  +\left( \frac{1}{-l_1^+ + \ii0} \right)
   \left( \frac{1}{ l_2^- + \ii0} \right)
  +\left( \frac{1}{ l_1^+ + \ii0} \right)
   \left( \frac{1}{ l_2^- + \ii0} \right)
  =
  -4\pi^2\delta(l_1^+)\delta(l_2^-).
  \label{eq:scet-spectator-same-cut-eikonal-sum}
\end{align}
The two spin-dependent spectator numerators contribute \(i^2=-1\), so the
same-side graphs enter with
\(+4\pi^2\delta(l_1^+)\delta(l_2^-)\).  With one gluon on each side of the
cut the eikonal sum has the opposite sign, but the spin phases are
\(i(-i)=+1\); hence those graphs also enter with
\(+4\pi^2\delta(l_1^+)\delta(l_2^-)\).  The dangerous Glauber region is
therefore not a phase that disappears in the measured double-spin cross
section.

The source calculation is the massless nonabelian spectator theory.  A
gauge-compatible way to regulate it before the leading-region reduction is to
work in \(d=4-2\epsilon\) and replace the lightlike eikonal directions by
Wilson lines along \(v_1(\rho)\) and \(v_2(\rho)\), with \(v_i^2<0\), taking
the lightlike limit only after the graph sum and subtractions have been
specified.  This off-lightcone regulator is a deformation of the Wilson-line
paths, not a gluon mass; it is a color scalar and therefore leaves the color
traces \(\operatorname{tr}(t^a)\) and
\(\operatorname{tr}(t^at^b)\operatorname{tr}(t^bt^a)\) unchanged.  It also
does not change the eikonal \(i0\) algebra above: the finite-\(\rho\)
denominators are \(v_i(\rho)\cdot l\pm\ii0\), and their leading Glauber
limit gives the same delta functions and the same same-side/opposite-side
relative signs.

In the double-spin channel, the separate-hadron leading-region subtraction is
explicitly
\[
\begin{aligned}
  \Delta\sigma_{\rm sep}^{(1,1)}
  &=
  H^{(0)}
  \bigl[\operatorname{tr}(t^a)\mathcal E_{1}^{a}\bigr]
  \bigl[\operatorname{tr}(t^b)\mathcal E_{2}^{b}\bigr]
  =0,\\
  \Delta\sigma_{\rm sep}^{(2,0)}+\Delta\sigma_{\rm sep}^{(0,2)}
  &=
  H^{(0)}
  \bigl[\Delta\Phi_1^{(2,ab)}\Phi_2^{(0)}
  +\Phi_1^{(0)}\Delta\Phi_2^{(2,ab)}\bigr]_{\rm double\ spin}
  =0 .
\end{aligned}
\]
The first term vanishes by
\eqref{eq:scet-spectator-single-gluon-trace-zero}; the second vanishes in the
double Sivers projection because the zeroth-order spectator trace has no
single-hadron transverse-spin asymmetry.  The cross-hadron graph sum is thus
not removable by a product of separate one-hadron objects.

After the on-shell spectator delta functions, the eikonal delta functions,
and these separate-hadron subtractions, the leading entangled source
contribution has the transverse skeleton
\begin{equation}
\begin{aligned}
  \Delta\sigma_{\rm ent}
  &=
  \frac{T_F^2(N_c^2-1)g^4}{s}
  \int
  \frac{\dd^2 k_{1\perp}}{(2\pi)^2}
  \frac{\dd^2 l_{1\perp}}{(2\pi)^2}
  \frac{\dd^2 l_{2\perp}}{(2\pi)^2}\,
  \mathcal H(k_1,k_3,k_4)
\\
  &\quad\times
  \frac{
  [\epsilon_{\perp jk}s_1^j l_1^k\,M_1(x_1)]
  [\epsilon_{\perp j'k'}s_2^{j'}l_2^{k'}\,M_2(x_2)]
  }{\mathcal D_{\rm spec}(k_{1\perp},l_{1\perp},l_{2\perp};q_\perp)}
\end{aligned}
  \label{eq:scet-spectator-entangled-breaking-integral}
\end{equation}
where
\[
\begin{aligned}
  \mathcal D_{\rm spec}
  &=
  l_{1\perp}^2\,
  l_{2\perp}^2\,
  [k_{1\perp}^2+L_1^2(x_1)]
\\
  &\quad\times
  [(q_\perp-k_{1\perp})^2+L_2^2(x_2)]\,
  [(l_{1\perp}-k_{1\perp})^2+L_1^2(x_1)]
\\
  &\quad\times
  [(l_{2\perp}-q_\perp+k_{1\perp})^2+L_2^2(x_2)] ,
\end{aligned}
\]
Here
\[
  M_i(x_i)=m_{H_i}(1-x_i)+m_{\psi_i},\qquad
  L_i^2(x_i)
  =
  (1-x_i)m_{q_i}^2+x_i m_{\psi_i}^2
  -x_i(1-x_i)m_{H_i}^2
\]
are the spin-dependent spectator numerator and the transverse mass function
of the Rogers--Mulders model.  The ellipses in the original displayed source
formula are therefore not arbitrary positive constants: for a chosen model
point they are fixed by \(x_i,m_{H_i},m_{q_i},m_{\psi_i}\).

The next finite integral is a sign witness for this reduced transverse
skeleton.  It is deliberately relabeled as an auxiliary positive-denominator
deformation, not as a regulated nonabelian gauge-theory calculation: replace
only the two reduced source factors \(l_{i\perp}^{-2}\) by
\((l_{i\perp}^2+\rho_i^2)^{-1}\) with \(\rho_i>0\), restrict
\(|k_{1x}|,|k_{1y}|\le3\), and keep the source-derived \(L_i^2\) and
\(M_i\).  This deformation is used only to obtain a finite, independently
checked, nonzero value of the already-identified entangled kernel.  Regulator
removal in the massless gauge theory and a normalized hadronic observable are
separate problems.

Integrated nonvanishing requires a sign or value statement over the whole
deformed domain.  The check uses the concrete source point
\[
\begin{gathered}
  x_1=x_2=\frac12,\quad
  m_{H_1}=1,\quad m_{H_2}=\frac65,\quad
  m_{q_1}=\frac34,\quad m_{q_2}=\frac45,\\
  m_{\psi_1}=\frac75,\quad m_{\psi_2}=\frac32,\quad
  M_X=2,\quad
  \rho_1=\rho_2=\frac13 ,
\end{gathered}
\]
so that
\[
  L_1^2=\frac{809}{800},\qquad
  L_2^2=\frac{217}{200},\qquad
  M_1=\frac{19}{10},\qquad
  M_2=\frac{21}{10}.
\]
For the hard scattering choose
\[
\begin{gathered}
  p_1^+=10,\quad p_2^-=\frac{37}{5},\quad
  k_3=\left(\frac52,\frac{49}{20},\frac72,0\right),\quad
  k_4=\left(\frac52,\frac54,-\frac52,0\right),\\
  q_\perp=k_{3\perp}+k_{4\perp}=(1,0),
  \qquad
  k_1=(5,0,k_x,k_y),\quad
  k_2=\left(0,\frac{37}{10},1-k_x,-k_y\right),
\end{gathered}
\]
where the ordered entries are \((+,-,x,y)\).  On the square
\(|k_x|,|k_y|\le3\), the Rogers--Mulders hard factor
\[
  \mathcal H_{\rm RM}(k_x,k_y)
  =
  \frac{1}{2x_1x_2s}
  \left[
    \frac{(k_1+k_3)\cdot(k_2+k_4)}
         {(k_1-k_3)^2-M_X^2}
  \right]^2
\]
obeys the explicit positive bound \(5.2\times10^{-3}\le
\mathcal H_{\rm RM}\le 3.3\times10^{-1}\).  The deformed transverse integral
over \(l\) obeys
\[
  \int\frac{\dd^2l}{(2\pi)^2}\,
  \frac{\epsilon_{\perp jk}s^j l^k}
       {[l_\perp^2+\rho^2][(l_\perp-K_\perp)^2+L^2]}
  =
  \frac{\epsilon_{\perp jk}s^jK^k}{4\pi}\,
  A_\rho(K_\perp^2;\rho^2,L^2),
\]
where
\[
  A_\rho(K^2;\rho^2,L^2)
  =
  \int_0^1
  \frac{(1-z)\,\dd z}
       {z\rho^2+(1-z)L^2+z(1-z)K^2}
  >0 .
\]
Thus the reduced integrand contains
\[
  [\epsilon_\perp(\hat x,k_{1\perp})]
  [\epsilon_\perp(\hat x,q_\perp-k_{1\perp})]
  =
  -k_{1y}^2 .
\]
All other factors in this deformed patch are positive.  The contribution is
therefore nonpositive pointwise and strictly negative on any transverse window
with nonzero \(k_{1y}\)-measure.  The local calculation establishes a nonzero
integrated obstruction in the positive-deformation witness for the displayed
double-spin projection; the physical factorization-breaking statement is the
source Rogers--Mulders graph-sum obstruction, not the auxiliary deformation by
itself.

The failure of \eqref{eq:scet-generalized-tmd-ansatz} is now sharp.  The
same-hadron two-gluon terms force any separate TMD repair to use
color-traced Wilson loops.  But a single gluon on such a loop gives
\(\operatorname{tr}(t^a)=0\), so the product of separate TMDs has no
order-\(g\) factor from \(H_1\) times an order-\(g\) factor from \(H_2\) that
could generate \eqref{eq:scet-spectator-entangled-breaking-integral}.  Adding
such a term by hand would spoil
\eqref{eq:scet-spectator-single-gluon-trace-zero}.  The missing object is a
joint color/Glauber operator entangling the two incoming hadrons, not a more
ornate Wilson line inside each separate TMD.

What remains possible is correspondingly narrower.  Inclusive collinear
factorization can still hold after the transverse recoil is integrated and
the relevant Glauber contributions cancel in the tested topology.
Color-singlet small-\(q_\perp\) TMD factorization, such as Drell--Yan with
the stated regulator, soft factor, \(Y\)-term, and Glauber proof obligations,
is not ruled out by this counterexample.  Process-specific joint Glauber
operators or nonfactorized color-density coordinates may describe the same
physics, but they are not the separate-hadron generalized TMD ansatz
\eqref{eq:scet-generalized-tmd-ansatz}.
\end{controlledapproximation}

\begin{hypothesis}[Leading-power SCET factorization structure]
\label{hyp:leading-power-scet-factorization}
A leading-power SCET factorization statement for an observable \(F\) consists
of a small parameter \(\lambda\), a list of modes with momentum scaling, a
mode-separated operator basis through the declared order in \(\lambda\), a
renormalized matrix-element definition of each hard, jet, beam, and soft
function, an overlap-subtraction prescription, a measurement decomposition on
those modes, a statement that Glauber or other factorization-breaking exchanges
either cancel or are represented by additional operators, and a remainder
estimate in powers of \(\lambda\) on the kinematic region where the formula is
used.  If the formula is asserted as a theorem, the estimate must be stated in
a topology on smeared cross sections or cumulants; otherwise it is a
controlled approximation hypothesis.
\end{hypothesis}

\begin{controlledapproximation}[Distributional leading-power factorization estimate]
\label{ca:scet-distributional-factorization-estimate}
The content of a leading-power factorization formula is an estimate on
smeared physical cross sections, not an equality of pointwise functions.
Assume that, for the chosen hard scale \(Q\), the physical QCD final-state
observable \(e\) determines a positive cross-section measure
\(\dd\mu_Q(e)\) on an endpoint window \(W_Q\).  Let
\(\mathcal B_{\rm end}\) be a Banach space of endpoint test functions
\(\varphi\colon W_Q\to\mathbb C\), with norm
\(\|\varphi\|_{\mathcal B_{\rm end}}\), strong enough to control the singular
distributions under discussion.  The physical smeared functional is
\begin{equation}
  \Sigma_Q^{\rm phys}(\varphi)
  =
  \int_{W_Q}\varphi(e)\,\dd\mu_Q(e).
  \label{eq:scet-physical-smeared-functional}
\end{equation}
A two-jet operator specification \(\mathcal D_{\rm SCET}\) as in
Definition~\ref{def:scet-two-jet-operator-datum} is treated, once the
specified regulator, operator existence, overlap subtraction, rapidity
subtraction when needed, and renormalization scheme have been supplied, as a
singular distributional coordinate
\(\dd\mu_{Q,\rm sing}^{\rm SCET}(e;\mathcal D_{\rm SCET})\).  In that
conditional coordinate system its smeared functional is
\begin{equation}
  \Sigma_{Q,\rm sing}^{\rm SCET}(\varphi;\mathcal D_{\rm SCET})
  =
  \left\langle
  H\otimes J_n\otimes J_{\bar n}\otimes S,\varphi
  \right\rangle_{\mathcal D_{\rm SCET}},
  \label{eq:scet-factorized-smeared-functional}
\end{equation}
where the pairing includes the measurement convolution, the endpoint
transform convention if one is used, and the subtraction prescription.

A controlled leading-power factorization claim on \(W_Q\) is the assertion
that there exist an exponent \(p>0\), a scale-dependent endpoint parameter
\(\lambda_W\), and a remainder bound \(R_Q\) such that
\begin{equation}
  \left|
  \Sigma_Q^{\rm phys}(\varphi)
  -
  \Sigma_{Q,\rm sing}^{\rm SCET}(\varphi;\mathcal D_{\rm SCET})
  \right|
  \le
  R_Q\,\|\varphi\|_{\mathcal B_{\rm end}},
  \qquad
  R_Q\le C_Q\,\lambda_W^{p}
  \label{eq:scet-distributional-factorization-bound}
\end{equation}
for every \(\varphi\in\mathcal B_{\rm end}\) in the declared kinematic
range.  If the hard, jet, soft, or anomalous-dimension entries are
perturbatively truncated, the displayed \(R_Q\) must be enlarged by the
declared perturbative truncation remainder in the same norm.  The common
phrases "leading power" and "factorization theorem" therefore have a definite
meaning only after the physical functional, test space, factorization structure,
endpoint parameter, and remainder topology have all been specified.

The same formulation also fixes the status of finite scheme changes.  In a
scalar transform-space chart, a finite multiplicative coordinate change
\begin{equation}
  H'=H R_H,\qquad
  \widetilde J_n'=\widetilde J_n R_n,\qquad
  \widetilde J_{\bar n}'=\widetilde J_{\bar n} R_{\bar n},\qquad
  \widetilde S'=\widetilde S R_S
  \label{eq:scet-finite-multiplicative-scheme-change}
\end{equation}
represents the same singular distribution only if
\begin{equation}
  R_H R_n R_{\bar n} R_S=1
  \label{eq:scet-finite-scheme-product-condition}
\end{equation}
as an identity in the transform variable and scale.  The anomalous dimensions
then transform as
\begin{equation}
  \gamma_i'
  =
  \gamma_i+\mu\frac{\dd}{\dd\mu}\log R_i,
  \qquad i=H,J_n,J_{\bar n},S,
  \label{eq:scet-finite-scheme-gamma-shift}
\end{equation}
and the sum of the transformed anomalous dimensions still vanishes because
\[
  \mu\frac{\dd}{\dd\mu}\log(R_HR_nR_{\bar n}R_S)=0 .
\]
In a mixing problem the same statement is a similarity-covariant statement
about matrix kernels and path-ordered products.  It is part of the coordinate
structure of the factorized approximation, not an optional convention added after
the formula has been written.
\end{controlledapproximation}

\paragraph{Reference map for factorization claims.}
The table below refines Definition~\ref{def:eft-prediction-datum} only for
reader navigation.  Each row starts from a physical observable, names the
regulated mathematical object used to approximate it, and then records the
strongest local status of the displayed formula.  It is not independent
evidence for any factorization theorem.  A named remainder is a placeholder
until the cited row supplies an estimate or a uniform bound in the stated
test-function topology.  A formula therefore inherits the status in its own row
unless a later local theorem supplies the missing regulator choice, Glauber
treatment, and remainder estimate.
A row is included when the manuscript makes a process-level hard, collinear,
soft, rapidity, Regge, or Abelian soft-factorization assertion for an
observable.  Closely adjacent identities are grouped only when they are
operator definitions, consistency conditions, or local algebra inside the same
observable-level row.  The table deliberately distinguishes those statements
from other homonyms: pole or residue factorization, large-\(N\)
expectation-value factorization, rank-one spinor or BCFW pole factorization,
causal factorization in perturbative algebraic QFT, BV
prefactorization-algebra terminology, instanton determinant or
coefficient-decomposition factorizations, finite-range Gaussian
decompositions, Monte Carlo conditional-probability factorizations,
conformal or categorical algebraic decompositions, supersymmetric coset or
Chern--Simons vector identities, quantum-spectral-curve gauge identities, and
integrable factorized scattering do not carry the same Glauber or
power-remainder content.

{\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{0.25\textwidth}>{\raggedright\arraybackslash}p{0.71\textwidth}@{}}
\hline
\textbf{Manuscript label / source} &
\textbf{Observable / regulator / source / status / Glauber / remainder}
\\ \hline
\endfirsthead
\hline
\textbf{Manuscript label / source} &
\textbf{Observable / regulator / source / status / Glauber / remainder}
\\ \hline
\endhead
\eqref{eq:jet-event-shape-factorization};
Definition~\ref{def:scet-two-jet-operator-datum};
Vol.~II Ch.~19b &
Observable: \(e^+e^-\) two-jet endpoint event shape.  Regulator: SCET
operator package with overlap and possible rapidity subtraction.  Source:
SCET construction lineage
\cite{BauerFlemingLuke2000SCET,BauerFlemingPirjolStewart2001SCET} plus the
local construction.  Proof: operator specification and conditional
leading-singular distribution.  Glauber: absent for the color-singlet
final-state two-jet observable
unless a later hadronic measurement introduces spectator exchange.  Remainder:
\(O(e^0)\) is nonsingular-distribution status; quantitative control is
supplied by \eqref{eq:scet-distributional-factorization-bound}.
\\ \hline
Hypothesis~\ref{hyp:leading-power-scet-factorization};
Controlled Approximation~\ref{ca:scet-distributional-factorization-estimate};
Vol.~II Ch.~19b &
Observable: generic leading-power SCET observable, specialized below by
channel.  Regulator: declared modes, operator definitions, measurement
factorization, overlap subtraction, and test-function topology.  Source: local
claim architecture.  Proof: this row is the proof obligation itself, not an
external theorem.  Glauber: a required choice, namely cancel, include a
Glauber operator, or fail.  Remainder: bounded by \(R_Q\) in the stated dual
norm, enlarged by perturbative truncation.
\\ \hline
\eqref{eq:scet-factorized-smeared-functional};
Vol.~II Ch.~19b &
Observable: the generic smeared SCET factorized candidate for a physical
functional \(W_Q\).  Regulator: declared endpoint test space, mode
operators, overlap subtraction, and scheme choices inherited from the
surrounding controlled approximation.  Source: local distributional SCET
architecture.  Proof: definition of the candidate distribution, not yet a
theorem about \(W_Q\).  Glauber: whatever the declared package includes or
cancels; omission is a proof defect for any channel that needs it.  Remainder:
no bound is supplied by the candidate alone; the next row supplies the dual
norm estimate and perturbative truncation term.
\\ \hline
Controlled Approximation~\ref{ca:scet-distributional-factorization-estimate};
\eqref{eq:scet-distributional-factorization-bound};
Vol.~II Ch.~19b &
Observable: smeared endpoint cross-section functional.  Regulator: Banach
test space and declared factorized distribution.  Source: local
distributional estimate.  Proof: definition of what a leading-power claim must
prove.  Glauber: explicit choice, namely cancel, include an operator, or fail.
Remainder: bounded by \(R_Q\) in the stated dual norm, enlarged by
perturbative truncation.
\\ \hline
\eqref{eq:scet-generalized-tmd-ansatz};
Controlled Approximation~\ref{ca:scet-spectator-tmd-color-entanglement};
\eqref{eq:scet-spectator-entangled-breaking-integral};
Vol.~II Ch.~19b &
Observable: high-\(p_T\) nearly back-to-back hadron production, double
transverse spin, fixed \(q_\perp\).  Regulator/status: massless nonabelian
spectator model regulated at graph level by dimensional regularization and
off-lightcone Wilson-line rapidity directions; the finite numerical witness is
an auxiliary positive-denominator deformation of the reduced transverse
skeleton, with source-derived spectator masses and a fixed transverse window,
not a gluon-mass regulator for the gauge theory.  Source: Rogers--Mulders
\cite{RogersMulders2010NoGeneralizedTMD}.  Proof: counterexample to the
separate-hadron generalized TMD ansatz
\eqref{eq:scet-generalized-tmd-ansatz}; locally strengthened by a
source-derived integrated sign check for the displayed double-spin projection.
Glauber: nonzero entangled Glauber/color term with
\(C_{\rm ent}=T_F^2(N_c^2-1)\).  Remainder: regulator-removal, full hard
model, and all-order operator-completion questions remain outside this
separate-hadron ansatz.
\\ \hline
\eqref{eq:finite-non-global-dipole-evolution};
Vol.~II Ch.~19b &
Observable: angular veto or gap soft radiation.  Regulator: finite angular
cells, large-\(N_c\), strong energy ordering.  Source: non-global logarithm
lineage \cite{DasguptaSalam2001NonGlobal,BanfiMarchesiniSmye2002NonGlobal}
and local finite model.  Proof: finite soft-dipole model for BMS-type
real--virtual soft radiation under a non-global measurement.  Glauber: no
Glauber exchange is represented by this row; super-leading or hadron-hadron
Glauber questions are separate coordinates in the relevant hard-process
factorization statement, not outputs of the finite BMS model.  Remainder:
finite-cell, large-\(N_c\), strong-ordering, and matching errors must be
carried separately from global Sudakov terms.
\\ \hline
Controlled Approximation~\ref{ca:massive-vector-sudakov-chart};
\eqref{eq:massive-vector-sudakov-region};
\eqref{eq:massive-vector-sudakov-area};
Vol.~II Ch.~19b &
Observable: one energetic Standard-Model charge line with a massive
electroweak vector unresolved inside a specified high-energy jet or charge
sector.  Regulator: physical vector mass \(M_V\), hard scale \(Q\), the
soft-collinear chart \(0<y<L_V,\;0<x<y/2\), and a declared representation
\(R\).  Source: local massive-vector Sudakov chart, organized as the
electroweak analogue of a hard/soft/collinear factor.  Proof: finite
phase-space area for the one-line virtual Sudakov factor, not a full
electroweak factorization theorem for an arbitrary measured process.
Glauber/charge: no QCD Glauber-cancellation claim; when weak charge is
resolved the factorized object is a charge-density matrix, not a scalar
single-line exponential.  Remainder: real-emission inclusiveness, weak-charge
channel sums, \(M_V/Q\) power terms, QCD jet/soft matching, resonance
approximations, and hadronization or track inputs must be listed separately.
\\ \hline
Definition~\ref{def:soft-drop-mass-factorization-datum};
\eqref{eq:soft-drop-mass-factorization-datum};
Vol.~II Ch.~19b &
Observable: soft-drop groomed jet mass.  Regulator: renormalized
hard/global-soft/jet/collinear-soft package with soft-drop scales.  Source:
soft-drop observable source \cite{LarkoskiMarzaniSoyezThaler2014SoftDrop}
plus local factorization input.  Proof: conditional factorization structure
for a specified measurement and channel.  Glauber: listed as an obstruction
when the measurement is not inclusive enough.  Remainder: \(R_{\rm SD}\)
records power, clustering, non-global, Glauber, and nonperturbative pieces.
\\ \hline
Controlled Approximation~\ref{ca:soft-drop-mass-factorized-resummation};
\eqref{eq:soft-drop-factorization-rg-consistency};
Vol.~II Ch.~19b &
Observable: perturbative resummation of the soft-drop groomed-mass singular
distribution.  Regulator: the same hard/global-soft/jet/collinear-soft scheme
plus perturbative scale profiles.  Source: local soft-drop factorization
structure and the soft-drop observable lineage
\cite{LarkoskiMarzaniSoyezThaler2014SoftDrop}.  Proof: controlled only when
all natural scales remain perturbative, the boundary functions and anomalous
dimensions are specified, and \(R_{\rm SD}\) is bounded.  The RG-consistency
condition is grouped here because it is the common-scale invariance test for
this same resummation structure, not a separate measured observable.  Glauber:
unchanged from the previous row.  Remainder: nonperturbative collinear-soft,
profile, matching, non-global, clustering, and power residuals.
\\ \hline
\eqref{eq:qcd-eec-back-to-back-factorization};
Controlled Approximation~\ref{ca:qcd-eec-tested-back-to-back-recoil};
Vol.~II Ch.~19 &
Observable: back-to-back EEC recoil chart.  Regulator: impact-parameter and
rapidity-renormalized TMD-like scheme.  Source: QCD
factorization/resummation framework \cite{CollinsSoperSterman1989Factorization}
plus local tested recoil chart.  Proof: conditional recoil factorization
hypothesis for a color-singlet source.  Glauber: must be absent or represented
in the recoil operator structure, not inferred from the cusp calculation.
Remainder:
\(q_T/Q\), large-\(b\), endpoint-atom, and matching residuals are separate.
\\ \hline
\eqref{eq:qcd-measured-eec-factorization-budget};
Controlled Approximation~\ref{ca:qcd-measured-eec-prediction};
Vol.~II Ch.~19 &
Observable: calorimetric or selected-track EEC.  Regulator: endpoint partition
with overlap subtraction and track/fragmentation lift.  Source: local
measured-observable budget.  Proof: prediction architecture.  Glauber:
inherited from each open chart; track lift does not prove cancellation.
Remainder: \(R_{\tau,Q}\), endpoint atoms, and nonperturbative
track/fragmentation inputs.
\\ \hline
Figure~\ref{fig:qcd-dis-ope-light-ray-factorization};
\eqref{eq:qcd-common-factorization-budget};
Vol.~II Ch.~19 &
Observable: common comparison map used to compare QCD factorization
statements against an exact color-singlet observable.  Regulator: the
process-specific finite regulator, Wilson-line geometry, subtraction,
rapidity, and continuum-limit data named by each later row.  Source: local
comparison map.  Status: reference map and residual
decomposition, not a theorem for any one process.  The DIS figure is grouped
with this row and the DIS row below because it records the operator-coordinate
source of the inclusive DIS factorization statement; it is not an independent
proof of a power bound.
Glauber: appears in the boundary term whenever the process probes it.
Remainder: all rows must place their residuals inside
\(\mathcal E_{\mathfrak P,Q}^{\rm fact}\) or state why the coordinate is only
structural.
\\ \hline
\eqref{eq:dis-factorization};
Hypothesis~\ref{hyp:qcd-dis-leading-twist-factorization};
Vol.~II Ch.~19 &
Observable: inclusive DIS on a compact \(x\)-window.  Regulator: renormalized
light-ray PDFs in a collinear factorization scheme.  Source: hard-process
factorization lineage \cite{CollinsSoperSterman1989Factorization} and local
hypothesis.  Proof: leading-twist hypothesis with displayed distributional
power bound.  Glauber: inclusive final-state sum is part of the standard proof
obligation; no TMD Glauber-sensitive recoil is measured.  Remainder:
compact-\(x\) higher-twist bound; threshold and small-\(x\) regions excluded.
\\ \hline
\eqref{eq:qcd-endpoint-kernel-cusp-coefficient};
\eqref{eq:qcd-large-spin-dglap-cusp};
\eqref{eq:qcd-dis-threshold-factorized-coordinate};
Vol.~II Ch.~19 &
Observable: DIS threshold boundary \(x_B\to1\) or large Mellin spin.  Regulator:
renormalized light-ray cusp/eikonal operator and plus-distribution scheme.
Source: local cusp and DGLAP derivation.  Proof: perturbative
operator-renormalization coordinate, not a full threshold factorization theorem.
Glauber: not tested by the cusp anomalous dimension alone.  Remainder: a
threshold theorem still needs hard, jet, soft, Mellin/Laplace inversion data,
and a joint \(Q,N\) power estimate.
\\ \hline
\eqref{eq:qcd-tmd-factorization};
Hypothesis~\ref{hyp:qcd-tmd-factorization-hypothesis};
Vol.~II Ch.~19 &
Observable: color-singlet small-\(q_\perp\) hadron-hadron production.
Regulator: rapidity-regulated TMDs with soft subtraction and \(Y\)-term.
Source: CSS/TMD lineage \cite{CollinsSoperSterman1989Factorization}.  Proof:
hypothesis until regulator, soft factor, and power topology are supplied.
Glauber: cancellation or operator representation is load-bearing.  Remainder:
\(q_\perp/Q\), \(Y\)-term, large-\(b\), and hadron-mass residuals.
\\ \hline
\eqref{eq:qcd-drell-yan-integrated-factorization};
Controlled Approximation~\ref{ca:qcd-drell-yan-integrated-theorem-boundary};
Vol.~II Ch.~19 &
Observable: integrated Drell--Yan in a compact \((Q^2,y)\) window.  Regulator:
integrated PDFs and tested Wightman functional.  Source: CSS hard-process
factorization lineage \cite{CollinsSoperSterman1989Factorization} plus local
proof-obligation map.  Proof: theorem boundary and tested conditional map;
the formula alone is a coordinate.  Glauber: inclusive-projection or
commutator bound required by the boundary list.  Remainder: the lead, soft,
Glauber, power, and regulator residuals are separately bounded in
Remark~\ref{rem:qcd-drell-yan-tested-factorization-proof-obligation-map}.
\\ \hline
\eqref{eq:qcd-drell-yan-tmd-factorization};
Hypothesis~\ref{hyp:qcd-drell-yan-tmd-factorization-hypothesis};
Vol.~II Ch.~19 &
Observable: Drell--Yan small-\(q_\perp\) spectrum.  Regulator: past-pointing
TMDs, rapidity scales, \(Y\)-term.  Source: CSS/TMD lineage
\cite{CollinsSoperSterman1989Factorization}.  Proof: conditional TMD
factorization hypothesis.  Glauber: explicit proof or controlled hypothesis
required.  Remainder: \(Y\)-term, \(q_\perp/Q\), hadron-mass, and
nonperturbative transverse-distance residuals.
\\ \hline
\eqref{eq:qcd-gpd-definition};
Section~\ref{sec:qcd-gpd-light-ray-matrix-elements};
Vol.~II Ch.~19 &
Observable: DVCS or related exclusive amplitudes using GPD coefficient
convolutions.  Regulator: renormalized off-forward light-ray matrix element in
a collinear scheme.  Source: local GPD definition and polynomiality.  Proof:
only the GPD matrix element and moment polynomiality are supplied here; the
exclusive-process factorization theorem is an additional leading-power
assertion.  Glauber: endpoint, soft-overlap, and spectator-exchange status must
be supplied by the process theorem.  Remainder: power, endpoint, skewness, and
subtraction residuals are outside the GPD definition.
\\ \hline
Definition~\ref{def:qcd-exclusive-pion-factorization-scheme};
\eqref{eq:qcd-pion-form-factor-factorization};
\eqref{eq:qcd-pion-transition-factorization};
Vol.~II Ch.~19 &
Observable: large-\(Q\) exclusive pion form factors.  Regulator:
leading-twist pion LCDA and hard coefficient in a common collinear scheme.
Source: exclusive factorization lineage
\cite{BrodskyLepage1980Exclusive,EfremovRadyushkin1980PionFormFactor}.
Proof: leading-power assertion with endpoint and soft qualifications.
Glauber: endpoint/soft overlap is the relevant obstruction; inclusive Glauber
cancellation is not automatic.  Remainder: \(O(\alpha_s,Q^{-4})_{\rm fac}\)
is scheme-bound and not uniform without endpoint control.
\\ \hline
\eqref{eq:qcd-small-x-leading-dis-dipole-functional};
Controlled Approximation~\ref{ca:qcd-small-x-leading-dis-dipole-channel};
Vol.~II Ch.~19 &
Observable: leading small-\(x\) DIS dipole channel.  Regulator:
rapidity-cutoff Wilson lines, impact factor, projective transverse regulator.
Source: dipole/JIMWLK lineage
\cite{NikolaevZakharov1991,Balitsky1996HighEnergy,Kovchegov1999SmallXF2}.
Proof: measured-observable controlled approximation.  Glauber: high-energy
eikonal Wilson lines encode the leading exchange; a separate TMD
Glauber-cancellation theorem remains an additional proof obligation.
Remainder: impact-factor, rapidity, projective, evolution, closure, endpoint,
and finite-\(Q\) residuals.
\\ \hline
\eqref{eq:track-function-test-pairing};
Controlled Approximation~\ref{ca:leading-endpoint-shape-function-convolution};
Vol.~II Ch.~19b &
Observable: track functions and endpoint shape-function convolution.
Regulator: renormalized lightlike operator distributions paired with tests.
Source: local operator datum and convolution estimate.  Proof:
nonperturbative input definitions plus conditional matching coordinates.
Glauber: inherited from the short-distance factorization used to match them.
Remainder: distributional pairing, matching, and moment/shape residuals must
be stated for the observable.
\\ \hline
Hypothesis~\ref{hyp:triple-regge-factorization-system};
\eqref{eq:triple-regge-factorization-formula};
Vol.~II Ch.~7 &
Observable: inclusive high-energy \(a b\to c X\) density in the triple-Regge
limit.  Regulator: complex angular-momentum continuation, signature convention,
six-point discontinuity, and background-growth bound.  Source: Regge
phenomenology organized locally as a hypothesis.  Proof: hypothesis, not a
derivation from local QFT.  Glauber: not a SCET Glauber-cancellation claim;
the exchanged high-energy Regge singularities are part of the assumed
discontinuity representation.  Remainder: background term bounded below the
retained trajectories on the stated domain.
\\ \hline
Theorem~\ref{thm:weinberg-leading-soft-photon};
\eqref{eq:weinberg-leading-soft-photon};
Figure~\ref{fig:volume-ii-soft-factorization};
Vol.~II Ch.~22 &
Observable: one added soft photon in the infrared-regulated hard kernel.
Regulator: fixed infrared regulator, physical polarization, and nonexceptional
massive charged hard data.  Source: Weinberg leading soft theorem plus the
local scalar/spinor external-line derivation.  Proof: leading eikonal
external-line poles and Ward-identity decoupling of a pure-gauge
polarization.  Glauber: no nonabelian Glauber coordinate.  Remainder:
amplitude-level \(O(\omega^0)\), not an inclusive-rate cancellation.
\\ \hline
\eqref{eq:qed-bloch-nordsieck-inclusive-factorization};
Figure~\ref{fig:volume-ii-bloch-nordsieck-cancellation};
Remark~\ref{rem:abelian-scope-bloch-nordsieck};
Vol.~II Ch.~22 &
Observable: Abelian inclusive transition probability with unresolved photons
below detector threshold \(E_T\).  Regulator: photon mass or soft energy
cutoff \(\mu\), with the degenerate final-state sum performed before removing
\(\mu\).  Source: local Bloch--Nordsieck calculation.  Proof: real
independent-emission exponent multiplied by the virtual soft factor at fixed
regulator, giving \eqref{eq:qed-bloch-nordsieck-inclusive-factorization}.
Glauber: no nonabelian Glauber coordinate; Coulomb phases, collinear
degeneracy, and dressed charged states are separate Abelian infrared data.
Remainder: subleading soft, collinear, finite-resolution, hard-matching, and
state-dressing terms.
\\ \hline
\end{longtable}
}

The most common excluded homonyms are fixed as follows.  The bound-state and
resonance pole-factorization statements, such as
Theorem~\ref{thm:bethe-salpeter-pole-factorization}, are residue statements at
isolated analytic singularities, not hard-region cross-section factorizations.
Large-\(N_c\) expectation-value factorization, such as
\eqref{eq:qcd-large-n-factorization}, is a connected-correlator suppression
statement.  Spinor-helicity rank-one factorization and tree pole residues, such
as \eqref{eq:null-momentum-spinor-factorization} and
\eqref{eq:tree-factorization-bcfw}, are on-shell analytic properties.  Causal
factorization in perturbative algebraic QFT and BV prefactorization algebras,
such as \eqref{eq:paqft-causal-factorization} and
Definition~\ref{def:bv-prefactorization-algebra}, are locality and
functoriality structures.  Algebraic-QFT split-property tensor
factorizations are operator-algebraic independence statements, not
hard-process soft/collinear decompositions.  Instanton determinant and coefficient
factorizations, finite-range Gaussian decompositions, Monte Carlo
conditional-probability factorizations, conformal and categorical algebraic
decompositions, supersymmetric coset or Chern--Simons vector identities,
quantum-spectral-curve gauge identities, and integrable factorized scattering
are likewise outside this hard-process reference map.  None of these rows is
assigned a QCD Glauber-cancellation or power-remainder coordinate merely
because it contains the word ``factorization''.

Renormalization of the hard, jet, and soft functions gives evolution
equations.  The resummation structure is clearest after a Laplace transform, or
any equivalent transform, has converted the measurement convolution into an
ordinary product.  Write the transformed singular coordinate as
\[
  \widetilde C(\nu,Q;\mu)
  =
  H(Q,\mu)\,
  \widetilde J_n(\nu,Q,\mu)\,
  \widetilde J_{\bar n}(\nu,Q,\mu)\,
  \widetilde S(\nu,Q,\mu),
\]
where \(\nu\) is conjugate to the event-shape measurement.  A scalar
renormalization-group transport structure consists of anomalous dimensions
\[
  \mu\frac{\dd}{\dd\mu}\log H=\gamma_H,\qquad
  \mu\frac{\dd}{\dd\mu}\log \widetilde J_a=\gamma_{J_a},\qquad
  \mu\frac{\dd}{\dd\mu}\log \widetilde S=\gamma_S ,
\]
all in the same subtraction and rapidity scheme, satisfying the consistency
condition
\begin{equation}
  \gamma_H+\gamma_{J_n}+\gamma_{J_{\bar n}}+\gamma_S=0 .
  \label{eq:scet-rg-consistency-transform-space}
\end{equation}
In a distributional formulation the same equation is an identity of
convolution kernels.  If color or operator mixing is present, the scalar
exponentials below are replaced by path-ordered evolution matrices and the
consistency condition is the corresponding statement that the product
observable has no residual common-scale dependence.

For a scalar transport structure, define
\[
  U_i(\mu,\mu_i)
  =
  \exp\!\left\{
    \int_{\log\mu_i}^{\log\mu}\gamma_i(\nu,Q;\mu')\,\dd\log\mu'
  \right\},
  \qquad
  i=H,J_n,J_{\bar n},S .
\]
The resummed singular coordinate is then
\begin{equation}
\begin{aligned}
  \widetilde C_{\rm sing}(\nu,Q)
  &=
  H(Q,\mu_H)U_H(\mu,\mu_H)\,
  \widetilde J_n(\nu,Q,\mu_J)U_{J_n}(\mu,\mu_J)
\\
  &\quad\times
  \widetilde J_{\bar n}(\nu,Q,\mu_J)U_{J_{\bar n}}(\mu,\mu_J)\,
  \widetilde S(\nu,Q,\mu_S)U_S(\mu,\mu_S).
  \label{eq:scet-resummed-transport-coordinate}
\end{aligned}
\end{equation}
Differentiating \(\log\widetilde C_{\rm sing}\) with respect to the common
scale \(\log\mu\) gives the left-hand side of
\eqref{eq:scet-rg-consistency-transform-space}; hence the expression is
independent of \(\mu\) inside the accuracy to which the anomalous dimensions
and matching functions have been supplied.  The natural choices
\(\mu_H\sim Q\), \(\mu_J\sim Q\sqrt e\), and \(\mu_S\sim Qe\), or their
transform-space analogues, are not extra physics assumptions.  They are
coordinate choices that keep the boundary functions free of the large
logarithms subsequently generated by the evolution kernels.
Finite subtraction-scheme changes may redistribute terms between the
anomalous dimensions and the boundary functions.  Such a redistribution is a
valid coordinate change only when the hard, jet, and soft entries are changed
together so that \eqref{eq:scet-rg-consistency-transform-space} and the
product coordinate \(\widetilde C_{\rm sing}\) are preserved to the declared
order.

If \(L\) denotes the large logarithm of the measured hierarchy, logarithmic
accuracy is a statement about a definite truncation of
\eqref{eq:scet-resummed-transport-coordinate}.  At LL one keeps the cusp
anomalous dimension and beta-function data needed to generate the leading
tower in the exponent, with tree-level boundary functions.  NLL adds the next
tower by including the next cusp data, the first non-cusp anomalous
dimensions, and the corresponding running-coupling terms.  NNLL adds the next
tower and the matching constants required by the declared convention.
Primed labels such as NLL\(^{\prime}\) or NNLL\(^{\prime}\) mean that fixed
boundary functions are included one order higher than the unprimed logarithmic
transport.  The labels are therefore incomplete unless the observable, the
factorization theorem, the transform convention, the natural scales, the
anomalous-dimension orders, the matching orders, and the matching prescription
to fixed order are all listed.  They are formal perturbative-accuracy
statements in a logarithmic hierarchy, not numerical error bounds once the
hierarchy is lost, a Landau-pole region is approached, or an unproved
factorization remainder dominates.

\subsection{Massive-Vector Sudakov Boundaries}

High-energy jet measurements in the Standard Model involve a second
short-distance boundary in addition to the QCD soft and collinear boundaries:
massive electroweak gauge bosons can be soft or collinear relative to an
energetic charged line while their physical masses stop the logarithmic
integration.  The correct object is a massive-vector factorization coordinate
for a specified observable and charge sector.  The physical mass \(M_V\) and
the hard scale \(Q\) are part of the electroweak perturbative datum, while
the QCD hadronic measurement remains a physical final-state observable as
above.

Let \(R\) be a representation of the gauge factor carried by an energetic
line, and define its quadratic charge by
\[
  T_R^aT_R^a=C_R\,\mathbf 1_R .
\]
Here \(T_R^a\) are the canonical generators after the gauge field has been
canonically normalized, so \(\alpha_V=g_V^2/(4\pi)\) is the coupling that
appears in the emission probability.  The corresponding convention is tied to
the Yang--Mills normalization of
Chapter~\ref{ch:vol2-classical-yang-mills-matter}: the path-integral gauge
field is written with kinetic term
\(-\frac1{4g_{\rm YM}^2}\operatorname{tr}F_{\mu\nu}F^{\mu\nu}\), and
canonical perturbative fields carry the coupling in the vertices.

\begin{controlledapproximation}[Single-line massive-vector Sudakov chart]
\label{ca:massive-vector-sudakov-chart}
Fix an energetic external charge in representation \(R\), a hard scale
\(Q\), a massive vector boson of mass \(M_V\), and a hierarchy
\(Q\gg M_V\).  In the strongly ordered soft-collinear chart, use the transverse
momentum \(k_\perp\) and energy fraction \(z\) of an unresolved virtual vector
boson as coordinates.  Put
\[
  L_V=\log\frac{Q^2}{M_V^2},\qquad
  y=\log\frac{Q^2}{k_\perp^2},\qquad
  x=\log\frac1z .
\]
The massive vector permits the logarithmic range
\begin{equation}
  0<y<L_V,\qquad 0<x<\frac{y}{2},
  \label{eq:massive-vector-sudakov-region}
\end{equation}
where the second inequality is the soft-collinear energy ordering
\(z>k_\perp/Q\).  Approximating the emission measure by
\((\alpha_V C_R/\pi)\,\dd x\,\dd y\), the one-line virtual Sudakov coordinate
is
\begin{equation}
  \Delta_R^{(V)}(Q,M_V)
  =
  \exp\!\left[
    -\frac{\alpha_V C_R}{\pi}\,
    A_V(L_V)
  \right],
  \qquad
  A_V(L_V)
  =
  \int_0^{L_V}\dd y\int_0^{y/2}\dd x
  =
  \frac14L_V^2 .
  \label{eq:massive-vector-sudakov-area}
\end{equation}
The approximation consists precisely of the eikonal numerator, the
soft-collinear phase-space chart, fixed coupling, one unresolved massive
vector, and a scalar charge channel.
\end{controlledapproximation}

The derivation of \eqref{eq:massive-vector-sudakov-area} is the finite
phase-space geometry.  The lower transverse-momentum boundary
\(k_\perp=M_V\) gives \(y=L_V\), the upper boundary \(k_\perp=Q\) gives
\(y=0\), and the condition that the emitted vector have energy larger than
its transverse momentum gives \(x<y/2\).  The area is therefore the triangle
of base \(L_V\) and height \(L_V/2\).  The exponent in
\eqref{eq:massive-vector-sudakov-area} is the virtual no-emission coordinate
in this chart.  A real-emission calculation cancels the same logarithmic
coordinate only for measurements that sum over the corresponding unresolved
massive-vector final states and over the charge channels connected by those
emissions.  In electroweak applications the incoming beam and the measured
final state often select weak charge.  The factorized object is then a matrix
in the external charge density space, with entries built from products of
generators \(T_i^a\) on the participating energetic lines.  A scalar
exponential such as \eqref{eq:massive-vector-sudakov-area} is the
one-line diagonal representative of this more general charge-transport
structure.

This is the clean way to place boosted and electroweak jet statements inside
the monograph.  A boosted massive-particle substructure analysis is a
measurement on stable hadronic and leptonic final states, together with a
declared resonance-pole or narrow-width approximation when an unstable
particle label such as \(W,Z,t\), or \(H\) is used.  The small parameters are
then listed explicitly, for example
\[
  \frac{m_{\rm obj}}{p_T},\qquad
  \frac{\Gamma_{\rm obj}}{m_{\rm obj}},\qquad
  \frac{M_W}{Q},\qquad
  \lambda_{\rm jet},
\]
as applicable to the chosen observable.  The perturbative electroweak
Sudakov transport, the QCD jet and soft functions, the resonance
approximation, and the nonperturbative hadronization or track coordinates are
separate entries in the prediction.  Their remainders have different meanings and
must be stated separately before a numerical prediction can be assigned a
claimed accuracy.

\begin{controlledapproximation}[Perturbative prediction for an IRC-safe jet observable]
A perturbative prediction for a jet observable consists of:
the renormalized fixed-order hard functions through a stated order in
\(\alpha_s\); the logarithmic accuracy of any resummation, such as LL, NLL,
NNLL, or beyond; the factorization theorem and its power corrections; the
renormalized PDFs or fragmentation functions when they enter; and a
prescription for matching resummed and fixed-order regions without double
counting.
\end{controlledapproximation}

\subsection{Global and Non-Global Soft Measurements}

A factorization statement for a small event shape is much simpler when the
measurement is global, meaning that every soft direction is measured with a
weight compatible with the same soft function.  Jet vetoes, gaps between
jets, and many substructure cuts are not global.  Then a soft gluon emitted
into an unmeasured angular region is not vetoed, but it changes the color
source that can radiate a later softer gluon into the measured region.  This
is the mathematical origin of non-global logarithms.  The point is not the
name; it is that the measurement algebra is no longer generated by a single
linear Wilson-line soft function.

The following finite model records the part of the argument that can be made
without invoking an unregulated continuum angular integral.  Fix finitely many
angular cells \(\mathcal C\) away from the hard Wilson-line directions.  Let
\(\mathcal M\subset\mathcal C\) be the measured, vetoed cells and
\(\mathcal U=\mathcal C\setminus\mathcal M\) the unmeasured cells.  For every
ordered dipole \((i,j)\), let \(K_{ij}^c\ge0\) be the cell-integrated eikonal
emission kernel into cell \(c\).  In a continuum soft-dipole approximation,
one representative regulated coordinate for this rate is
\[
  K_{ij}^c
  =
  \int_{\Omega_c}\frac{\dd\Omega_k}{4\pi}\,
  \frac{\alpha_s C_{\rm dip}}{\pi}
  \frac{1-\widehat n_i\cdot\widehat n_j}
       {(1-\widehat n_i\cdot\widehat k)
        (1-\widehat k\cdot\widehat n_j)}
\]
after the collinear neighborhoods of \(i\) and \(j\) and the ultraviolet
coupling convention have been specified.  In the finite model the numbers
\(K_{ij}^c\) are simply positive regulator coordinates.

Let \(G_{ij}(L)\) denote the finite soft-dipole gap coordinate for the dipole
\((i,j)\), where \(L\) is the logarithm of the resolved soft-energy hierarchy,
and impose the initial condition \(G_{ij}(0)=1\).  The finite non-global
evolution equation is
\begin{equation}
  \frac{\dd}{\dd L}G_{ij}(L)
  =
  -A_{ij}G_{ij}(L)
  +
  \sum_{u\in\mathcal U}
  K_{ij}^{u}
  \bigl(G_{iu}(L)G_{uj}(L)-G_{ij}(L)\bigr),
  \qquad
  A_{ij}:=\sum_{m\in\mathcal M}K_{ij}^{m}.
  \label{eq:finite-non-global-dipole-evolution}
\end{equation}
The first term is direct veto radiation into the measured cells.  The second
term is real--virtual balanced for radiation into an unmeasured cell: if every
gap coordinate is identically one, it contributes zero.  Its nonlinear
product is the essential non-global structure: after an unmeasured emission
at direction \(u\), the original dipole is replaced by the two dipoles
\((i,u)\) and \((u,j)\), each of which can subsequently radiate into
\(\mathcal M\).

Differentiating \eqref{eq:finite-non-global-dipole-evolution} at \(L=0\)
gives a useful diagnostic.  Since \(G_{ij}(0)=1\),
\[
  G'_{ij}(0)=-A_{ij}.
\]
A second differentiation gives
\[
  G''_{ij}(0)
  =
  A_{ij}^2
  -
  \sum_{u\in\mathcal U}
  K_{ij}^{u}\bigl(A_{iu}+A_{uj}-A_{ij}\bigr).
\]
Therefore
\begin{equation}
  G_{ij}(L)
  =
  1-A_{ij}L
  +
  \frac{L^2}{2}
  \left[
    A_{ij}^2-\mathcal N_{ij}
  \right]
  +O(L^3),
  \qquad
  \mathcal N_{ij}
  :=
  \sum_{u\in\mathcal U}
  K_{ij}^{u}\bigl(A_{iu}+A_{uj}-A_{ij}\bigr).
  \label{eq:finite-non-global-second-order}
\end{equation}
The \(A_{ij}^2/2\) term is the square of the direct global Sudakov veto.  The
term \(\mathcal N_{ij}\) is the first finite non-global coefficient.  It
vanishes when there is no unmeasured region, and it also vanishes in the
special additive case \(A_{iu}+A_{uj}=A_{ij}\) for every unmeasured cell
\(u\).  A generic angular veto does not satisfy this identity.  Thus the
failure of globalness is not a cosmetic complication in logarithmic counting;
it changes the evolution problem from a linear anomalous dimension to a
nonlinear Wilson-line measurement problem.

The same calculation exposes the error made by a global-Sudakov-only ansatz.
If one replaces the measured soft function by
\[
  G_{ij}^{\rm glob}(L)=\exp(-A_{ij}L),
\]
then the first derivative at the origin is correct, but
\[
  G_{ij}^{\rm glob}(L)-G_{ij}(L)
  =
  \frac{L^2}{2}\mathcal N_{ij}+O(L^3).
\]
The residual is not a convention for distributing logarithms among hard, jet,
and soft factors.  It is the finite weight of histories in which a resolved
gluon first enters an unmeasured cell \(u\) and then the softer radiation from
the daughter dipoles \((i,u)\) or \((u,j)\) reaches the vetoed region.  This
unmeasured-to-measured cascade is the physical history that the global ansatz
has projected out.  The path decomposition
\[
  \mathcal N_{ij}
  =
  \sum_{u\in\mathcal U}
  \underbrace{K_{ij}^u}_{\text{unmeasured first emission}}
  \underbrace{(A_{iu}+A_{uj}-A_{ij})}_{\text{new veto rate minus virtual subtraction}}
\]
is therefore an observable-level obstruction to replacing the angular veto by
a product of independent global Sudakov factors.  In the additive boundary
case the bracket vanishes cell by cell, so the obstruction disappears for a
physical reason rather than by a choice of scale or scheme.

Equation~\eqref{eq:finite-non-global-dipole-evolution} is not a theorem about
full QCD.  It is a finite large-\(N_c\), strongly energy-ordered,
soft-dipole coordinate system for one obstruction to naive factorization.
Beyond this finite model one must specify the color density matrix at finite
\(N_c\), the regulator for collinear neighborhoods of the hard lines, the
matching to the hard and collinear functions, the treatment of recoil, and
the status of Glauber exchanges.  Super-leading logarithms, when present, are
not explained by the finite non-global coefficient alone; they signal an
additional factorization obstruction that must be stated in the operator
structure of Hypothesis~\ref{hyp:leading-power-scet-factorization}.

\section{Jet Substructure Observables and Grooming}

Jet substructure observables are measurements on the constituents of a
previously identified jet.  The jet definition, recombination scheme, grooming
algorithm, and substructure measurement are all part of the observable.  Their
IRC behavior must therefore be checked as a combined map from the final-state
calorimetric measure to a number or distribution.

\begin{definition}[Soft drop]
\label{def:soft-drop-grooming}
Start with the constituents of a jet of radius \(R_0\), recluster them with
the Cambridge--Aachen metric, and recursively decluster the last clustering
step into two branches \(i,j\).  Let
\[
  z_{ij}=\frac{\min(p_{Ti},p_{Tj})}{p_{Ti}+p_{Tj}},
  \qquad
  \theta_{ij}=\frac{\Delta R_{ij}}{R_0}.
\]
For parameters \(z_{\rm cut}>0\) and \(\beta_{\rm SD}\geq0\), the soft-drop
condition is
\[
  z_{ij}>z_{\rm cut}\,\theta_{ij}^{\beta_{\rm SD}} .
\]
If the condition fails, the softer branch is removed and the recursion
continues on the harder branch.  If the condition holds, the remaining
constituents define the groomed jet.  The case \(\beta_{\rm SD}=0\) is the
modified mass-drop tagger.
\end{definition}

\paragraph{IRC classification of soft-drop outputs.}
With an IRC-safe input jet algorithm and a finite transverse-momentum
threshold for the original jet, the following statements hold away from the
boundary hypersurfaces of the measurement.
\begin{enumerate}
  \item If \(\beta_{\rm SD}>0\), the soft-drop groomed four-momentum is IRC
  safe.
  \item If \(\beta_{\rm SD}=0\), the groomed four-momentum is not collinear safe
  as a standalone observable.
  \item For \(\beta_{\rm SD}=0\), any claimed IRC safety statement must name
  the measured functional of the groomed constituents, for example a groomed
  jet mass or an energy-correlation shape, and must check the collinear limit
  of that functional rather than the groomed four-vector itself.
\end{enumerate}

Add a particle of transverse momentum \(\lambda p_T\) with
\(\lambda\downarrow0\).  If it is tested against a hard branch at finite
angle, then \(z_{ij}=O(\lambda)\), while
\(z_{\rm cut}\theta_{ij}^{\beta_{\rm SD}}\) has a positive lower bound on
compact subsets away from \(\theta_{ij}=0\).  The soft branch is removed and
cannot change the groomed four-momentum except on the codimension-one
surface where the inequality is exactly saturated.  If the soft particle is
collinear to a hard branch, Cambridge--Aachen recombines or declusters it at
small \(\theta\), and its contribution to the branch momentum vanishes with
\(\lambda\).

For a collinear splitting of a hard particle, the two branches have
\(\theta_{ij}\downarrow0\).  If \(\beta_{\rm SD}>0\), then
\[
  z_{\rm cut}\theta_{ij}^{\beta_{\rm SD}}\longrightarrow0.
\]
For every fixed splitting fraction \(0<z<1\), the soft-drop condition is
therefore eventually satisfied, and the collinear pair is retained as the
parent four-momentum.  This gives the collinear limit of the groomed
four-vector.

If \(\beta_{\rm SD}=0\), the right-hand side of the soft-drop condition is the
fixed number \(z_{\rm cut}\).  Consider a massless one-particle jet of
four-momentum \(p\), and replace it by a collinear pair \(zp\) and
\((1-z)p\) with \(0<z<z_{\rm cut}<1/2\).  Cambridge--Aachen declustering sees
the two branches, the soft-drop condition fails, and the softer branch is
removed.  The groomed four-vector is \((1-z)p\), not \(p\), even in the exact
collinear limit.  Hence the groomed four-vector is not collinear safe for
\(\beta_{\rm SD}=0\).

This counterexample does not by itself classify every functional of the
groomed constituents.  A groomed mass, energy-correlation ratio, or tagged
substructure variable can still be IRC safe, but its safety is a statement
about that functional and its normalization.  The check must be made at the
level of Definition~\ref{def:irc-safe-measurement}.

\paragraph{Finite-tree formulation and decision margins.}
The preceding definition hides a useful finite mathematical structure: before one takes
continuum, perturbative, or detector limits, soft drop is a deterministic map
on a finite Cambridge--Aachen declustering tree.  Let \(T\) be a rooted binary
tree whose leaves are the constituents of a fixed input jet.  For every
internal vertex \(v\), let \(v_L,v_R\) be the two branches obtained by undoing
one Cambridge--Aachen recombination.  Write
\[
  E_{v,a}=\sum_{\ell\in v_a}p_{T\ell},
  \qquad
  z_v=\frac{\min(E_{v,L},E_{v,R})}{E_{v,L}+E_{v,R}},
  \qquad
  \theta_v=\frac{\Delta R(v_L,v_R)}{R_0},
\]
where the branch direction entering \(\Delta R(v_L,v_R)\) is the
four-momentum recombination direction of that branch.  A fixed-tree soft-drop
specification consists of \(T\), the leaf momenta, a declared tie-breaking rule when
\(E_{v,L}=E_{v,R}\), and the parameters \((z_{\rm cut},\beta_{\rm SD})\).
The recursion starts at the root.  At a vertex \(v\), if
\[
  z_v>z_{\rm cut}\theta_v^{\beta_{\rm SD}},
\]
the algorithm stops and retains precisely the leaves below \(v\).  If the
condition fails, it moves to the harder child and deletes the softer child.

This finite-tree specification is the object whose soft and collinear limits must be
tested before it is inserted into a partonic phase-space integral.  Define the
soft-drop decision margin and, on failed vertices, the harder-branch margin by
\begin{equation}
  m_v^{\rm SD}
  =
  \left|z_v-z_{\rm cut}\theta_v^{\beta_{\rm SD}}\right|,
  \qquad
  m_v^{\rm hard}
  =
  \frac{|E_{v,L}-E_{v,R}|}{E_{v,L}+E_{v,R}} .
  \label{eq:soft-drop-decision-margins}
\end{equation}
For a finite event with no zero-energy branch, all quantities \(z_v\) and
\(\theta_v\) are continuous functions of the leaf momenta on any chart where
the Cambridge--Aachen tree \(T\) itself is unchanged.  Hence, if every vertex
visited by the recursion has \(m_v^{\rm SD}>0\), and every failed vertex has
\(m_v^{\rm hard}>0\), there is a neighborhood of the leaf momenta in which
the same vertices are visited, the same branches are deleted, and the same
leaf set is retained.  If the tree \(T\) is generated from the momenta rather
than supplied as part of the input, one must additionally require positive
Cambridge--Aachen clustering margins: every pairwise angular distance
comparison that determines the tree must remain separated from its nearest
competitor.  These margins are the precise finite-event meaning of the phrase
``away from soft-drop and clustering boundaries.''  At such a point any
groomed observable that is a smooth function of the retained momenta has the
same local soft or collinear behavior as the retained-momentum map specified
by the fixed recursion.  The boundary hypersurfaces
\(m_v^{\rm SD}=0\), \(m_v^{\rm hard}=0\), and the Cambridge--Aachen distance
tie hypersurfaces require separate distributional analysis and cannot be
silently discarded when a singular measurement or endpoint resummation is
being defined.

\paragraph{Measured mMDT functionals.}
The \(\beta_{\rm SD}=0\) counterexample above is deliberately a counterexample
to the groomed four-vector.  It is not a license to classify every measured
functional of the groomed set by the same sentence.  A useful finite-event
test is the following.  For a retained finite leaf set \(S\), define the
homogeneous angular moment
\begin{equation}
  G_\kappa(S)
  =
  \frac{1}{E_S^2R_0^\kappa}
  \sum_{\ell,m\in S}
  E_\ell E_m\,\theta_{\ell m}^{\kappa},
  \qquad
  E_S=\sum_{\ell\in S}E_\ell,
  \qquad
  \kappa>0,
  \label{eq:groomed-angular-moment-finite}
\end{equation}
where \(\theta_{\ell m}\) is the angular distance and the diagonal terms
vanish.  In the one-prong collinear split used in the mMDT counterexample,
with energies \((1-z)E\) and \(zE\), \(0<z<z_{\rm cut}\), the mMDT recursion
retains only the harder daughter.  Therefore
\[
  G_\kappa(\{(1-z)E\})=0=G_\kappa(\{E\})
\]
in the exact collinear limit.  If \(\beta_{\rm SD}>0\), the same pair is
retained for sufficiently small opening angle \(\theta\), and
\[
  G_\kappa(\{(1-z)E,zE\})
  =
  2z(1-z)\left(\frac{\theta}{R_0}\right)^\kappa ,
\]
which tends to zero as \(\theta\downarrow0\).  Thus a positive-angular-power
shape such as \eqref{eq:groomed-angular-moment-finite} passes this elementary
collinear test even when the groomed four-vector does not.

By contrast, the retained-energy fraction
\[
  r_E(S)=\frac{E_S}{E_{\rm input}}
\]
is not collinear safe for mMDT as a standalone observable: the same exact
collinear split changes \(r_E\) from \(1\) to \(1-z\).  This is the
measurement-level lesson.  For \(\beta_{\rm SD}=0\), a groomed observable is
acceptable only after its own collinear boundary behavior is checked.  Angular
weights with positive powers can kill the deleted one-prong collinear branch;
energy-retention observables cannot.  Endpoint distributions of
\(G_\kappa\), groomed masses, or ratios built from them still require their
own factorization and resummation data, because IRC safety of the measurement
is only the first finiteness condition.

\paragraph{Fixed-coupling soft-drop mass radiator.}
A useful measured-observable calibration is the groomed jet-mass cumulative
distribution in the single-emission soft-collinear chart.  The calculation
below is the logarithmic phase-space geometry that a
factorization theorem must reproduce in its leading coordinate.

\begin{controlledapproximation}[Soft-drop mass at fixed-coupling soft-collinear accuracy]
\label{ca:soft-drop-mass-fixed-coupling-radiator}
Consider one energetic parton of color factor \(C_i\) and one unresolved
emission with energy fraction \(z\ll1\), normalized angle
\(\theta=\Delta R/R_0\ll1\), and groomed mass coordinate
\[
  \rho=\frac{m_{J,{\rm gr}}^2}{p_T^2R_0^2}=z\theta^2
\]
in this chart.  Approximate the real-emission measure by
\[
  \frac{\alpha_s C_i}{\pi}\frac{\dd\theta^2}{\theta^2}\frac{\dd z}{z},
\]
hold \(\alpha_s\) fixed, ignore finite terms in the splitting function,
recoil, boundary corrections at \(z\sim1\), clustering logarithms, non-global
correlations, and power corrections in \(\rho,z_{\rm cut}\).  The soft-drop
condition is \(z>z_{\rm cut}\theta^{\beta_{\rm SD}}\).  With
\[
  L_\rho=\log\frac1\rho,\qquad
  L_z=\log\frac1{z_{\rm cut}},\qquad
  b=\frac{\beta_{\rm SD}}{2},
\]
the leading radiator is the area
\begin{equation}
  A_\beta(L_\rho,L_z)
  =
  \int_0^\infty\dd t\int_0^\infty\dd u\,
  \mathbf 1_{\{u+t<L_\rho\}}\,
  \mathbf 1_{\{u<L_z+bt\}},
  \label{eq:soft-drop-mass-radiator-area}
\end{equation}
where \(t=\log(1/\theta^2)\) and \(u=\log(1/z)\).  Hence
\begin{equation}
  \Sigma_{\rm SD}^{\rm LL}(\rho)
  =
  \exp\!\left[
    -\frac{\alpha_s C_i}{\pi}A_\beta(L_\rho,L_z)
  \right]
  \label{eq:soft-drop-mass-fixed-coupling-sudakov}
\end{equation}
inside this controlled approximation.
\end{controlledapproximation}

The area in \eqref{eq:soft-drop-mass-radiator-area} is elementary but
instructive.  If \(L_\rho\leq L_z\), the mass constraint is stronger than the
grooming constraint throughout the triangular region \(u+t<L_\rho\), so
\[
  A_\beta(L_\rho,L_z)=\frac12 L_\rho^2 .
\]
If \(L_\rho>L_z\), the two boundary lines
\[
  u=L_\rho-t,\qquad u=L_z+bt
\]
intersect at
\[
  t_*=\frac{L_\rho-L_z}{1+b}.
\]
The area is therefore
\begin{equation}
  A_\beta(L_\rho,L_z)
  =
  L_z t_*+\frac{b}{2}t_*^2
  +\frac12(L_\rho-t_*)^2,
  \qquad
  t_*=\frac{L_\rho-L_z}{1+b},
  \qquad
  L_\rho>L_z .
  \label{eq:soft-drop-mass-radiator-closed-form}
\end{equation}
For mMDT, \(\beta_{\rm SD}=0\), this reduces to
\begin{equation}
  A_0(L_\rho,L_z)
  =
  L_\rho L_z-\frac12L_z^2,
  \qquad L_\rho>L_z .
  \label{eq:mmdt-single-log-radiator}
\end{equation}
Thus mMDT removes the soft double logarithm below the grooming scale: the
radiator grows linearly in \(L_\rho\), with coefficient fixed by
\(L_z=\log(1/z_{\rm cut})\).  For \(\beta_{\rm SD}>0\), the coefficient of
the asymptotic \(L_\rho^2\) term is nonzero and depends continuously on
\(\beta_{\rm SD}\); the grooming boundary still removes part of the soft
triangle, but no longer removes the double-logarithmic growth completely.

Equations
\eqref{eq:soft-drop-mass-fixed-coupling-sudakov}--\eqref{eq:mmdt-single-log-radiator}
should be read as a calibration of logarithmic phase space.  A physical
prediction for a groomed jet-mass distribution must replace this chart by a
renormalized factorization structure: the hard production process, the jet
function for the initiating parton, the groomed soft or collinear-soft
function, overlap subtractions, running coupling, finite splitting functions,
matching to fixed order, non-global and clustering effects where present, and
the hadronization coordinate appropriate to the measured mass region.

\paragraph{Renormalized soft-drop mass factorization structure.}
The preceding radiator becomes a QCD prediction only after the measured
groomed-mass coordinate has been embedded in a renormalized operator
definition.
Let \(Q\) denote the hard energy scale of the measured jet, for instance
\(Q=p_T R_0\) in a small-radius hadron-collider chart, and let
\[
  \rho=\frac{m_{J,{\rm gr}}^2}{Q^2}.
\]
In the soft-collinear coordinates of
Controlled Approximation~\ref{ca:soft-drop-mass-fixed-coupling-radiator}, the
emission that lies simultaneously on the mass boundary and on the soft-drop
boundary satisfies
\[
  z\theta^2=\rho,
  \qquad
  z=z_{\rm cut}\theta^{\beta_{\rm SD}}.
\]
Solving these two equations gives the boundary scales
\begin{equation}
  \theta_*=
  \left(\frac{\rho}{z_{\rm cut}}\right)^{1/(2+\beta_{\rm SD})},
  \qquad
  z_*=
  z_{\rm cut}^{2/(2+\beta_{\rm SD})}
  \rho^{\beta_{\rm SD}/(2+\beta_{\rm SD})}.
  \label{eq:soft-drop-boundary-scales}
\end{equation}
Thus \(z_*\theta_*^2=\rho\) and \(z_*=z_{\rm cut}\theta_*^{\beta_{\rm SD}}\).
These identities are not a factorization theorem, but they determine the
scales that any such theorem must see: ordinary collinear radiation carrying
the measured jet mass has virtuality of order \(Q^2\rho\), while radiation
near the grooming boundary has light-cone components of order
\[
  p_{\rm cs}^-\sim Qz_*,
  \qquad
  p_{\rm cs}^+\sim Q\rho,
  \qquad
  |p_{{\rm cs}\perp}|\sim Q\sqrt{\rho z_*}.
\]
Consequently its invariant transverse scale is
\begin{equation}
  \mu_{\rm cs}\sim Q\sqrt{\rho z_*},
  \qquad
  \mu_J\sim Q\sqrt{\rho},
  \qquad
  \mu_G\sim Qz_{\rm cut},
  \qquad
  \mu_H\sim Q .
  \label{eq:soft-drop-natural-scales}
\end{equation}
For \(\beta_{\rm SD}=0\), \(z_*=z_{\rm cut}\) and
\(\mu_{\rm cs}\sim Q\sqrt{\rho z_{\rm cut}}\); this is the scale separation
behind the single-logarithmic mMDT radiator in
\eqref{eq:mmdt-single-log-radiator}.

\begin{definition}[Soft-drop groomed-mass factorization structure]
\label{def:soft-drop-mass-factorization-datum}
Fix the jet algorithm, recoil convention, recombination scheme, hard process,
partonic channel \(i\), subtraction scheme, and soft-drop parameters
\((z_{\rm cut},\beta_{\rm SD})\).  A leading singular soft-drop groomed-mass
factorization structure consists of:
\begin{enumerate}
  \item a hard coefficient \(H_i(Q,\mu)\) for the short-distance production of
  the measured energetic parton or color representation;
  \item a global-soft veto coordinate \(S^{G}_i(z_{\rm cut},R_0,\mu)\)
  describing wide-angle soft radiation that is removed by grooming but affects
  the probability that the event lies in the selected groomed-jet sample;
  \item a renormalized collinear jet distribution \(J_i(\rho_J,\mu)\) and a
  renormalized collinear-soft distribution
  \(S^{\rm cs}_{i,{\rm SD}}(\rho_{\rm cs};z_{\rm cut},\beta_{\rm SD},R_0,\mu)\),
  both understood as distributions paired with endpoint test functions;
  \item the measurement decomposition
  \(\rho=\rho_J+\rho_{\rm cs}\) in the leading singular chart, together with
  the overlap-subtraction and rapidity-renormalization data needed to avoid
  double counting the common soft-collinear boundary;
  \item anomalous dimensions, in the same scheme, satisfying the common-scale
  consistency condition
  \begin{equation}
    \gamma_H+\gamma_G+\gamma_J+\gamma_{\rm cs}=0
    \label{eq:soft-drop-factorization-rg-consistency}
  \end{equation}
  as an identity of transform-space functions or, before transform, of
  convolution kernels;
  \item a remainder functional \(R_{\rm SD}\) on the stated test-function space,
  recording power corrections in \(\rho,z_{\rm cut},R_0\), finite splitting
  functions beyond the declared singular order, recoil and clustering effects,
  non-global correlations not included in \(S^G_i\), Glauber obstructions if
  the measurement is not inclusive enough, and nonperturbative coordinates when
  one of the scales in \eqref{eq:soft-drop-natural-scales} approaches
  \(\Lambda_{\rm QCD}\).
\end{enumerate}
The corresponding singular coordinate is
\begin{equation}
\begin{aligned}
  \left\langle
  \frac{\dd\Sigma^{\rm SD}_{i,{\rm sing}}}{\dd\rho},
  \varphi
  \right\rangle
  &=
  H_i(Q,\mu)\,S^G_i(z_{\rm cut},R_0,\mu)
  \int \dd\rho_J\,\dd\rho_{\rm cs}\,
  J_i(\rho_J,\mu)
  S^{\rm cs}_{i,{\rm SD}}(\rho_{\rm cs},\mu)
  \varphi(\rho_J+\rho_{\rm cs})
\\
  &\quad+
  \left\langle R_{\rm SD},\varphi\right\rangle ,
  \label{eq:soft-drop-mass-factorization-datum}
\end{aligned}
\end{equation}
for every endpoint test function \(\varphi\) in the declared domain.  The
symbols \(S^G_i\), \(J_i\), and \(S^{\rm cs}_{i,{\rm SD}}\) do not have
standalone physical meaning outside this common structure; finite scheme changes
may move terms among them only if
\eqref{eq:soft-drop-factorization-rg-consistency} and the paired distribution
\eqref{eq:soft-drop-mass-factorization-datum} are preserved.
\end{definition}

\begin{controlledapproximation}[Perturbative soft-drop mass resummation]
\label{ca:soft-drop-mass-factorized-resummation}
A perturbative resummation based on
Definition~\ref{def:soft-drop-mass-factorization-datum} is controlled on a
region of endpoint tests only when the scales
\(\mu_H,\mu_G,\mu_J,\mu_{\rm cs}\) in
\eqref{eq:soft-drop-natural-scales} remain in the perturbative regime, the
factorization remainder \(R_{\rm SD}\) is bounded in the declared topology, and
the logarithmic-accuracy label specifies the anomalous dimensions, boundary
matching functions, fixed-order matching prescription, rapidity scheme if
present, and scale-profile map used to avoid crossing nonperturbative
regions.  If \(\mu_{\rm cs}\) is of order \(\Lambda_{\rm QCD}\), the
collinear-soft distribution in
\eqref{eq:soft-drop-mass-factorization-datum} must be replaced by a
nonperturbative shape coordinate or by a regulator-level reconstruction
observable with its own matching theorem.  A Monte-Carlo tune or an empirical
model may supply a useful phenomenological coordinate, but it is then an
additional input, not a derivation of the soft-drop mass distribution from
perturbative QCD.
\end{controlledapproximation}

\paragraph{Groomed momentum sharing and Sudakov-conditioned coordinates.}
A second soft-drop observable, often denoted \(z_g\), records the momentum
sharing at the first declustering step that passes the soft-drop condition.
For a finite Cambridge--Aachen tree in the notation above, let \(v_\ast\) be
the first vertex visited by the recursion for which
\[
  z_{v_\ast}>z_{\rm cut}\theta_{v_\ast}^{\beta_{\rm SD}} .
\]
If such a vertex exists, define
\[
  z_g=z_{v_\ast},
  \qquad
  \theta_g=\theta_{v_\ast}.
\]
If no such vertex exists, the pair \((z_g,\theta_g)\) is not a number attached
to the event; one must either work with the event-selection indicator or
declare a separate null value outside the numerical range.  This elementary
point is already enough to show that \(z_g\) is not an ordinary IRC-safe
measurement in the sense of Definition~\ref{def:irc-safe-measurement}: in the
one-particle collinear limit there is no accepted split, while a nearby
two-particle configuration may have a finite accepted value.  The useful
short-distance statement about \(z_g\) is therefore not fixed-order IRC
safety.  It is an all-order, Sudakov-conditioned statement inside a specified
strongly ordered soft-collinear approximation.

\begin{controlledapproximation}[Sudakov-conditioned groomed sharing]
\label{ca:sudakov-conditioned-zg}
Fix an initiating parton channel \(i\), a normalized angle
\(\vartheta=\theta/R_0\in(0,1]\), a soft-collinear ordering in decreasing
\(\vartheta\), and a positive symmetrized splitting kernel
\(\overline P_i(z)\) on \(0<z\leq1/2\) in the declared subtraction scheme.
In the fixed-coupling, independent-emission, strongly ordered chart, the
accepted-emission rate is
\begin{equation}
  \dd\Gamma_i
  =
  a_i\,\overline P_i(z)\,
  \frac{\dd\vartheta}{\vartheta}\,\dd z\,
  \mathbf 1_{\{z>z_{\rm cut}\vartheta^{\beta_{\rm SD}}\}},
  \qquad
  a_i=\frac{\alpha_s C_i}{\pi}.
  \label{eq:zg-accepted-emission-rate}
\end{equation}
Let
\begin{equation}
  \mathcal P_i(\zeta)
  =
  \int_{\zeta}^{1/2}\overline P_i(z)\,\dd z,
  \qquad
  \mathcal R_i(\vartheta)
  =
  a_i\int_{\vartheta}^{1}
  \frac{\dd\vartheta'}{\vartheta'}\,
  \mathcal P_i\!\left(z_{\rm cut}{\vartheta'}^{\beta_{\rm SD}}\right),
  \qquad
  \Delta_i(\vartheta)=e^{-\mathcal R_i(\vartheta)} .
  \label{eq:zg-sudakov-factor}
\end{equation}
Here \(\Delta_i(\vartheta)\) is the probability, inside this chart, that no
emission at an angle larger than \(\vartheta\) has passed grooming.  The joint
density of the first accepted split is then
\begin{equation}
  \dd\mathbb P_i(z_g,\vartheta_g)
  =
  \Delta_i(\vartheta_g)\,
  a_i\,\overline P_i(z_g)\,
  \frac{\dd\vartheta_g}{\vartheta_g}\,\dd z_g\,
  \mathbf 1_{\{z_g>z_{\rm cut}\vartheta_g^{\beta_{\rm SD}}\}} .
  \label{eq:zg-first-accepted-joint-density}
\end{equation}
This is a controlled approximation only after the kernel, coupling
prescription, ordering variable, recoil convention, endpoint regularization,
and matching to the rest of the factorization structure have been specified.
\end{controlledapproximation}

The derivation of
\eqref{eq:zg-first-accepted-joint-density} is the stopping-time identity for
the ordered emission process.  The probability of no accepted split before
angle \(\vartheta_g\) is \(\Delta_i(\vartheta_g)\).  The probability of one
accepted split in the infinitesimal cell
\((\vartheta_g,\vartheta_g+\dd\vartheta_g)\times
(z_g,z_g+\dd z_g)\) is the rate
\eqref{eq:zg-accepted-emission-rate}.  Multiple emissions in the same
infinitesimal ordered cell are higher order in the cell volume.  Multiplying
these two factors gives
\eqref{eq:zg-first-accepted-joint-density}.  This argument does not replace a
QCD factorization theorem; it is the finite probabilistic core that such a
theorem must reproduce in its leading logarithmic coordinate.

The special case \(\beta_{\rm SD}=0\) exposes the mechanism transparently.
Then \(\mathcal P_i(z_{\rm cut}\vartheta^{\beta_{\rm SD}})
=\mathcal P_i(z_{\rm cut})\) is independent of \(\vartheta\), so
\[
  \Delta_i(\vartheta)=
  \vartheta^{a_i\mathcal P_i(z_{\rm cut})}.
\]
If the idealized chart is allowed to run to arbitrarily small angle, the
probability of eventually finding an accepted split is one, and integrating
\eqref{eq:zg-first-accepted-joint-density} over \(\vartheta_g\) gives
\begin{equation}
  p_i^{\rm mMDT}(z_g)
  =
  \frac{\overline P_i(z_g)}
       {\displaystyle\int_{z_{\rm cut}}^{1/2}
        \overline P_i(z)\,\dd z}\,
  \mathbf 1_{\{z_{\rm cut}<z_g\leq1/2\}} .
  \label{eq:mmdt-zg-leading-distribution}
\end{equation}
Indeed,
\[
  a_i\overline P_i(z_g)
  \int_0^1
  \vartheta^{a_i\mathcal P_i(z_{\rm cut})}
  \frac{\dd\vartheta}{\vartheta}
  =
  \frac{\overline P_i(z_g)}{\mathcal P_i(z_{\rm cut})}.
\]
With a finite angular cutoff \(\vartheta_{\min}\), the same expression is the
conditional density after dividing by the accepted-event probability
\(1-\Delta_i(\vartheta_{\min})\).  The cancellation of \(a_i\) in
\eqref{eq:mmdt-zg-leading-distribution} is not a statement that the
observable is nonperturbative-free; it is a consequence of conditioning on the
first accepted splitting inside this leading ordered chart.  Running coupling,
finite splitting functions, flavor changes, hadron masses, detector
thresholds, and nonperturbative collinear dynamics all belong to the
factorization and remainder data, not to the stopping-time identity alone.

For \(\beta_{\rm SD}>0\), a value \(z_g<z_{\rm cut}\) can occur only at an
angle satisfying
\[
  \vartheta_g<\left(\frac{z_g}{z_{\rm cut}}\right)^{1/\beta_{\rm SD}} .
\]
The fixed-order expansion of the \(z_g\) distribution is then singular at the
collinear boundary, while the all-order expression
\eqref{eq:zg-first-accepted-joint-density} may still define a finite
distribution because \(\Delta_i(\vartheta)\) suppresses the small-angle
region.  This is the precise sense, within the stated approximation, in which
the coordinate is Sudakov-conditioned.  It is a useful perturbative object
only together with the Sudakov factor, the conditioning convention, and the
remainder estimate for the measured jet sample.

\begin{definition}[\(N\)-subjettiness]
\label{def:n-subjettiness}
For a jet with constituents \(k\), choose \(N\) axes
\(\{a_1,\ldots,a_N\}\) by a specified IRC-safe axis prescription.  For
\(\beta_\tau>0\),
\[
  \tau_N^{(\beta_\tau)}
  =
  \frac{1}{p_{TJ}R_0^{\beta_\tau}}
  \sum_{k\in J}p_{Tk}\,
  \min_{a=1,\ldots,N}(\Delta R_{ka})^{\beta_\tau}.
\]
Ratios such as \(\tau_{21}=\tau_2/\tau_1\) are substructure coordinates only
on the domain where the denominator is separated from zero; near the
denominator boundary their distribution requires the same factorization and
resummation data as the underlying \(\tau_N\).
\end{definition}

The axis prescription is part of the observable.  One prescription with a
clean nonperturbative formulation is global minimization over axes.  Let
\((Y,d)\) be the compact angular patch of a jet, with
\(\operatorname{diam}Y\le R_0\), and let \(\mu\) be the positive transverse
energy measure of the jet constituents on \(Y\).  For \(M=\mu(Y)>0\), define
\begin{equation}
  \tau_{N,\min}^{(\beta_\tau)}(\mu)
  =
  \frac1{M R_0^{\beta_\tau}}
  \inf_{(a_1,\ldots,a_N)\in Y^N}
  \int_Y
  \min_{1\le r\le N}d(\mathbf n,a_r)^{\beta_\tau}\,
  \dd\mu(\mathbf n).
  \label{eq:minimized-n-subjettiness-measure}
\end{equation}
For an event consisting of finitely many hadrons this is the usual
\(N\)-subjettiness value with axes chosen by global minimization.  The set of
minimizing axes can have several components at symmetric configurations; the
observable in \eqref{eq:minimized-n-subjettiness-measure} is the minimized
value, not a chosen point in the minimizer set.

\begin{lemma}[Continuity of minimized \(N\)-subjettiness]
\label{lem:minimized-subjettiness-continuity}
Fix \(N\ge1\), \(\beta_\tau>0\), \(0<m_0<E<\infty\), and restrict to positive
measures \(\mu\) on \(Y\) with \(m_0\le \mu(Y)\le E\).  The functional
\(\tau_{N,\min}^{(\beta_\tau)}\) of
\eqref{eq:minimized-n-subjettiness-measure} is weakly continuous.  If a soft
constituent of transverse energy \(\varepsilon\) is added, then
\begin{equation}
  \left|
  \tau_{N,\min}^{(\beta_\tau)}(\mu+\varepsilon\delta_{\mathbf m})
  -
  \tau_{N,\min}^{(\beta_\tau)}(\mu)
  \right|
  \le \frac{2\varepsilon}{m_0}.
  \label{eq:minimized-subjettiness-soft-bound}
\end{equation}
If two constituents of total transverse energy \(E_{ab}\) and angular
separation \(\theta\) are replaced by their transverse-energy parent direction
\(\mathbf n_{ab}\), and if
\[
  d(\mathbf n_a,\mathbf n_{ab})+d(\mathbf n_b,\mathbf n_{ab})
  \le c\,\theta ,
\]
then
\begin{equation}
  \left|
  \Delta\tau_{N,\min}^{(\beta_\tau)}
  \right|
  \le
  \frac{E_{ab}}{m_0R_0^{\beta_\tau}}\,
  C_{\beta_\tau,R_0}\,(c\theta)^{\alpha_\tau},
  \qquad
  \alpha_\tau=\min(\beta_\tau,1),
  \label{eq:minimized-subjettiness-collinear-bound}
\end{equation}
where one may take \(C_{\beta,R_0}=1\) for \(0<\beta\le1\) and
\(C_{\beta,R_0}=\beta R_0^{\beta-1}\) for \(\beta\ge1\).
\end{lemma}

\begin{proof}
For \(A=(a_1,\ldots,a_N)\in Y^N\), set
\[
  \Phi_A(\mathbf n)=\min_{1\le r\le N}d(\mathbf n,a_r)^{\beta_\tau}.
\]
The space \(Y^N\) is compact, and \(A\mapsto\Phi_A\) is continuous as a map
from \(Y^N\) to \(C(Y)\) with the uniform norm.  Therefore the infimum in
\eqref{eq:minimized-n-subjettiness-measure} is attained.  The family
\(\{\Phi_A:A\in Y^N\}\) is compact in \(C(Y)\).  If \(\mu_j\to\mu\) weakly
with uniformly bounded total mass, then
\[
  \sup_{A\in Y^N}
  \left|
  \int\Phi_A\,\dd\mu_j-\int\Phi_A\,\dd\mu
  \right|\longrightarrow0 .
\]
Indeed, approximate the compact family \(\{\Phi_A\}\) by a finite
\(\eta\)-net in \(C(Y)\), use weak convergence on the finite net, and then
let \(\eta\downarrow0\).  Taking infima over \(A\) is continuous under this
uniform convergence, and division by \(M=\mu(Y)\), bounded away from zero,
gives weak continuity of \(\tau_{N,\min}^{(\beta_\tau)}\).

For a soft addition, write
\[
  I(\mu)=\inf_A\int\Phi_A\,\dd\mu .
\]
Since \(0\le\Phi_A\le R_0^{\beta_\tau}\), one has
\[
  |I(\mu+\varepsilon\delta_{\mathbf m})-I(\mu)|
  \le \varepsilon R_0^{\beta_\tau}.
\]
Also \(I(\mu)\le M R_0^{\beta_\tau}\).  With
\(M' = M+\varepsilon\),
\[
  \left|
  \frac{I(\mu+\varepsilon\delta_{\mathbf m})}{M'R_0^{\beta_\tau}}
  -
  \frac{I(\mu)}{MR_0^{\beta_\tau}}
  \right|
  \le
  \frac{\varepsilon}{M'}+
  \frac{\varepsilon}{M'}
  \le \frac{2\varepsilon}{m_0}.
\]

For collinear recombination, the total mass \(M\) is unchanged.  The function
\(t\mapsto t^{\beta_\tau}\) on \([0,R_0]\) is
\(\alpha_\tau\)-Hölder with the displayed constant:
\[
  |u^{\beta_\tau}-v^{\beta_\tau}|
  \le C_{\beta_\tau,R_0}|u-v|^{\alpha_\tau}.
\]
For every axis tuple \(A\),
\[
  |\Phi_A(\mathbf n_a)-\Phi_A(\mathbf n_{ab})|
  +
  |\Phi_A(\mathbf n_b)-\Phi_A(\mathbf n_{ab})|
  \le C_{\beta_\tau,R_0}(c\theta)^{\alpha_\tau},
\]
up to replacing \(c\) by a harmless larger constant if the parent-direction
bound is stated separately for each daughter.  Hence the change in
\(\int\Phi_A\,\dd\mu\) is bounded by
\(E_{ab}C_{\beta_\tau,R_0}(c\theta)^{\alpha_\tau}\), uniformly in \(A\).
Taking infima before and after recombination and dividing by
\(MR_0^{\beta_\tau}\), with \(M\ge m_0\), gives
\eqref{eq:minimized-subjettiness-collinear-bound}.
\end{proof}

Thus minimizing-axis \(N\)-subjettiness is an IRC-continuous calorimetric
observable as a value functional.  An algorithm that returns one selected
minimizing axis may jump at degenerate events, but the minimized value used
in the weighted cross section is continuous in the soft and collinear senses
above.  Ratios such as \(\tau_{21}\) inherit this control only on regions
where the denominator is bounded away from zero; otherwise their endpoint
distribution is a separate singular observable.

Energy-correlation functions, defined in
Chapter~\ref{ch:volume-ii-19}, provide a detector-based substructure
alternative.  They depend only on the positive calorimetric measure and
angular kernels, whereas \(N\)-subjettiness also depends on an axis-finding
map.  Both are useful, but their mathematical inputs differ.

\section{Energy Correlators as Calorimetric Coordinates}
\label{sec:energy-correlators-calorimetric-coordinates}

The cleanest way to formulate jet substructure without assigning physical
status to partons is to regard an event as a positive energy measure on the
celestial sphere or on a jet patch.  For a final state \(X\), write
\[
  \mu_X=\sum_{r\in X}E_r\,\delta_{\mathbf n_r},
  \qquad
  \mathbf n_r=\frac{\mathbf p_r}{|\mathbf p_r|},
\]
or, inside a jet of reference energy \(E_{\rm ref}\),
\[
  \bar\mu_X=\frac{1}{E_{\rm ref}}\mu_X .
\]
The total mass of \(\bar\mu_X\) is one for a fully normalized jet and is at
most one if acceptance cuts or grooming remove energy.  A calorimetric
observable is a function, distribution, or positive functional of this
measure.  The energy-flow detector of
Chapter~\ref{ch:volume-ii-19} is the operator whose spectral value on a
hadronic out-state is precisely \(\mu_X\).

\begin{definition}[Smeared energy-correlator functional]
\label{def:smeared-energy-correlator-functional}
Let \(K\in C((S^2)^k)\).  The \(K\)-smeared \(k\)-point energy-correlator
functional of a finite positive Borel measure \(\mu\) on \(S^2\) is
\begin{equation}
  \mathcal C_K^{(k)}(\mu)
  =
  \int_{(S^2)^k}
  K(\mathbf n_1,\ldots,\mathbf n_k)\,
  \dd\mu(\mathbf n_1)\cdots\dd\mu(\mathbf n_k).
  \label{eq:smeared-energy-correlator-functional}
\end{equation}
If \(K\) is supported away from coincidence diagonals, this is a
separated-angle detector product.  If \(K\) meets a diagonal, the functional
includes the contact part in which two detector insertions act on the same
hadron.  Both are part of the detector observable; removing a contact term is
a different distributional convention.
\end{definition}

The diagonal convention carries more information before angular projection
than after it.  Pushing the two-detector product forward by
\(\zeta=\mathbf n_1\cdot\mathbf n_2\) sends every same-hadron diagonal point
to the endpoint atom at \(\zeta=1\), with total mass
\(\sum_{r\in X}E_r^2\).  A finite calorimeter product instead retains the
diagonal measure \(\sum_{r\in X}E_r^2\,
\delta_{(\mathbf n_r,\mathbf n_r)}\) on \((S^2)^2\).  Moving the same total
diagonal mass from one detector cell to another leaves the projected
\(\zeta=1\) atom, the zeroth EEC moment, and the first EEC moment unchanged,
while it changes a same-cell product
\(\mathcal C^{(2)}_{\mathbf 1_B\otimes \mathbf 1_B}(\mu_X)\).  Comparisons
with binned response or covariance data therefore transport contact atoms
with their detector-cell coordinates before the final opening-angle
pushforward.

The usual EEC is obtained by choosing a sequence of kernels
\[
  K_\Delta(\mathbf n_1,\mathbf n_2;\chi)
  \longrightarrow
  \delta(\cos\chi-\mathbf n_1\cdot\mathbf n_2)
\]
as angular distributions.  Energy-correlation functions used for
substructure, such as \eqref{eq:qcd-energy-correlation-function-jet}, are
special cases with kernels built from powers of pairwise angular distances
and with diagonal terms omitted or killed by positive powers of the
distance.

\begin{lemma}[Continuity of smeared energy correlators]
\label{lem:smeared-energy-correlator-continuity}
Let \(K\) be Lipschitz on \((S^2)^k\), and restrict to measures with total
mass at most \(E\).  Then \(\mathcal C_K^{(k)}\) is continuous in the
weak topology of finite measures.  Moreover, if a final state is changed by
adding a soft particle of energy \(\varepsilon\), then
\begin{equation}
  |\Delta\mathcal C_K^{(k)}|
  \le
  k\,\|K\|_\infty\,E^{k-1}\varepsilon
  +O(\varepsilon^2).
  \label{eq:energy-correlator-soft-bound}
\end{equation}
If two particles with total energy \(E_{ab}\) and angular separation
\(\theta_{ab}\) are replaced by their energy-weighted parent direction, then
\begin{equation}
  |\Delta\mathcal C_K^{(k)}|
  \le
  k\,\operatorname{Lip}(K)\,E^{k-1}E_{ab}\,O(\theta_{ab})
  \label{eq:energy-correlator-collinear-bound}
\end{equation}
on compact regions where the parent direction is defined.
\end{lemma}

\begin{proof}
Weak continuity follows because \(K\) is continuous on the compact space
\((S^2)^k\), and product measures \(\mu_j^{\otimes k}\) converge weakly to
\(\mu^{\otimes k}\) when \(\mu_j\to\mu\) weakly with uniformly bounded total
mass.  This can be checked first on factorized kernels
\(K=f_1\otimes\cdots\otimes f_k\), where the statement is the product of
\(\int f_i\,\dd\mu_j\to\int f_i\,\dd\mu\), and then extended to all
continuous \(K\) by the Stone--Weierstrass theorem on \((S^2)^k\).

For a soft addition, write
\(\mu'=\mu+\varepsilon\delta_{\mathbf m}\).  Expanding
\((\mu+\varepsilon\delta_{\mathbf m})^{\otimes k}-\mu^{\otimes k}\), the
terms linear in \(\varepsilon\) insert the new point in one of \(k\) slots
and integrate the other \(k-1\) slots against \(\mu\).  Their absolute value
is bounded by \(k\|K\|_\infty E^{k-1}\varepsilon\); the remaining terms have
at least two factors of \(\varepsilon\).

For a collinear recombination, the signed measure difference is
\[
  \delta\mu
  =
  E_a\delta_{\mathbf n_a}
  +E_b\delta_{\mathbf n_b}
  -E_{ab}\delta_{\mathbf n_{ab}},
  \qquad E_{ab}=E_a+E_b .
\]
It has total mass zero.  In each tensor slot, pairing \(\delta\mu\) with the
Lipschitz function obtained by fixing the other \(k-1\) angles gives a bound
by \(\operatorname{Lip}(K)E_{ab}O(\theta_{ab})\).  Summing over the \(k\)
possible slots and bounding the other integrations by \(E^{k-1}\) gives
\eqref{eq:energy-correlator-collinear-bound}.  In terms with two or more
insertions of \(\delta\mu\), choose one insertion and use its zero total mass
against the Lipschitz dependence on that slot; the remaining
\(\delta\mu\)-insertions are bounded by their total variation, at most
\(2E_{ab}\).  Thus every such term still carries the same
\(O(\theta_{ab})\) factor, with a constant depending on \(k\), \(E\),
\(E_{ab}\), and \(\|K\|_\infty+\operatorname{Lip}(K)\).  This proves the
displayed IRC-continuity estimates.
\end{proof}

\begin{theorem}[Energy-correlator polynomials separate calorimetric observables]
\label{thm:energy-correlator-polynomial-density}
Let
\[
  \mathcal M_E(S^2)=\{\mu\ge0:\mu(S^2)\le E\}
\]
be the compact space of finite positive Borel measures on \(S^2\), equipped
with the weak topology.  Let \(\mathcal A_E\) be the real algebra generated
by the linear energy moments
\[
  \mu\longmapsto \int_{S^2}f(\mathbf n)\,\dd\mu(\mathbf n),
  \qquad f\in C(S^2).
\]
Equivalently, \(\mathcal A_E\) is the algebra of finite linear combinations
of smeared energy correlators
\(\mathcal C_K^{(k)}\) with factorized kernels
\(K=f_1\otimes\cdots\otimes f_k\).  Then \(\mathcal A_E\) is uniformly dense
in \(C(\mathcal M_E(S^2))\).
\end{theorem}

\begin{proof}
The space \(\mathcal M_E(S^2)\) is compact in the weak topology.  Indeed,
positive measures with total mass at most \(E\) are a weak-\(*\) closed
subset of the radius-\(E\) ball in the dual of \(C(S^2)\), and that ball is
weak-\(*\) compact by Banach--Alaoglu; because \(C(S^2)\) is separable, this
compact topology is metrizable.  The algebra \(\mathcal A_E\) contains
constants by taking the empty product, and it is closed under multiplication
because products of linear moments are generated by the algebra definition.

It separates points.  If \(\mu\ne\nu\), the signed measure \(\mu-\nu\) is a
nonzero continuous linear functional on \(C(S^2)\).  By the Riesz
representation theorem there is an \(f\in C(S^2)\) such that
\[
  \int f\,\dd\mu\ne\int f\,\dd\nu .
\]
Thus the linear moment \(M_f(\mu)=\int f\,\dd\mu\) distinguishes \(\mu\) and
\(\nu\).  The real Stone--Weierstrass theorem applies to the compact
Hausdorff space \(\mathcal M_E(S^2)\): a real subalgebra of
\(C(\mathcal M_E(S^2))\) that contains constants and separates points is
uniformly dense.  Hence \(\mathcal A_E\) is dense.

A monomial
\[
  M_{f_1}(\mu)\cdots M_{f_k}(\mu)
  =
  \int_{(S^2)^k}
  f_1(\mathbf n_1)\cdots f_k(\mathbf n_k)\,
  \dd\mu(\mathbf n_1)\cdots\dd\mu(\mathbf n_k)
\]
is exactly \(\mathcal C_K^{(k)}(\mu)\) for the factorized kernel
\(K=f_1\otimes\cdots\otimes f_k\), by the defining property of the product
measure \(\mu^{\otimes k}\).  Finite linear combinations of such monomials
are therefore exactly the finite factorized-kernel energy-correlator
polynomials, so the dense algebra is an energy-correlator algebra.
\end{proof}

The theorem is a finite-resolution statement.  It applies to continuous
functions of the calorimetric measure on the compact energy ball
\(\mathcal M_E(S^2)\).  Sharp step-function cuts, exact jet-counting
boundaries, and distributional angular kernels are obtained as limits of
continuous observables, and their boundary terms must be analyzed separately.
Graph-indexed energy-flow polynomials used in jet physics choose a concrete
family of angular kernels built from pairwise distances.  Treating that
smaller family as a practical basis requires an additional approximation
statement for angular kernels on the detector resolution and symmetry class
being used.  The object-level reason for their robustness is the theorem
above: energy correlators are coordinates on the positive energy measure
itself, not on a partonic history.

\begin{controlledapproximation}[Finite-resolution energy-correlator observable assembly]
\label{ca:finite-resolution-eec-observable-assembly}
Fix an angular metric \(d\) on \(S^2\), a finite detector partition
\(\mathcal P_\delta=\{U_a\}\), representatives \(\mathbf n_a\in U_a\),
and a diameter bound \(\operatorname{diam}(U_a)\leq\delta\).  For a positive
finite-energy event measure \(\mu\), set
\begin{equation}
  \pi_\delta\mu
  =
  \sum_a \mu(U_a)\,\delta_{\mathbf n_a}.
  \label{eq:finite-resolution-eec-binned-measure}
\end{equation}
Let \(K\in C((S^2)^k)\) be Lipschitz with constant \(L_K\) for the product
metric \(d_k((\mathbf n_i),(\mathbf m_i))=\sum_{i=1}^k d(\mathbf n_i,\mathbf
m_i)\), and assume \(\mu(S^2)\leq E\).  Then
\begin{equation}
  \left|
  \mathcal C_K^{(k)}(\mu)
  -
  \mathcal C_K^{(k)}(\pi_\delta\mu)
  \right|
  \leq
  k L_K E^k\delta .
  \label{eq:finite-resolution-eec-lipschitz-bound}
\end{equation}
Indeed, couple each \(\mathbf n\in U_a\) to its representative
\(\mathbf n_a\) and replace the \(k\) detector slots one at a time.  Each
replacement changes the kernel by at most \(L_K\delta\), while the remaining
energy integrations contribute at most \(E^k\).

A physical prediction for an energy-correlator bin therefore requires more
data than the coordinate theorem.  It consists of:
\begin{enumerate}
  \item the binned detector measure \(\pi_\delta\mu\), including the chosen
  convention for diagonal detector contacts and endpoint atoms when a sharp
  EEC distribution is represented by limits of smooth kernels;
  \item a short-distance light-ray or hard/jet/soft coefficient chart in a
  specified ultraviolet, rapidity, and endpoint subtraction scheme;
  \item the nonperturbative lift required by the measurement, such as a shape,
  fragmentation, track, or hadronization coordinate rather than an implicit
  partonic final state;
  \item an observable-error decomposition
  \[
    B_{\rm bin}+B_{\rm contact}+B_{\rm pert}
    +B_{\rm FG}+B_{\rm np}+B_{\rm track}+B_{\rm cont}
  \]
  in which \(B_{\rm bin}\) is the proved finite-resolution estimate from
  \eqref{eq:finite-resolution-eec-lipschitz-bound}, \(B_{\rm contact}\)
  records the chosen contact and endpoint convention, \(B_{\rm pert}\) is a
  perturbative truncation estimate, \(B_{\rm FG}\) is the stated
  factorization or Glauber hypothesis residual, \(B_{\rm np}\) and
  \(B_{\rm track}\) are nonperturbative input uncertainties, and
  \(B_{\rm cont}\) is the continuum/regulator reconstruction residual.
\end{enumerate}
The assembled claim is a tested functional inequality: for every detector
test \(\varphi\) in the declared test space,
\begin{equation}
  \begin{aligned}
    \left|
    \left\langle O_{\rm phys}-O_{\rm pred},\varphi\right\rangle
    \right|
    &\leq
    B_{\rm bin}(\varphi)+B_{\rm contact}(\varphi)+B_{\rm pert}(\varphi)\\
    &\quad
    +B_{\rm FG}(\varphi)+B_{\rm np}(\varphi)
    +B_{\rm track}(\varphi)+B_{\rm cont}(\varphi).
  \end{aligned}
  \label{eq:finite-resolution-eec-observable-budget}
\end{equation}
The contact and endpoint terms are coordinates of the measured distribution,
with operational consequences: moving a small-angle annulus into an endpoint
atom, or replacing a track-selected diagonal by an average charged fraction,
changes the tested functional unless the compensating coordinate is
transported with it.  The role of
Theorem~\ref{thm:energy-correlator-polynomial-density} is to identify complete
finite-resolution calorimetric coordinates; the role of the short-distance and
nonperturbative entries above is to make a QCD prediction for those
coordinates.
\end{controlledapproximation}

\begin{controlledapproximation}[Measured small-angle EEC endpoint chart]
\label{ca:measured-small-angle-eec-endpoint-chart}
The small-angle endpoint of a measured EEC bin is the place where the
calorimetric coordinate theorem meets the light-ray OPE data of
Chapter~\ref{ch:volume-ii-19} and
Chapter~\ref{ch:cft-light-ray-energy-correlators}.  It is useful to state the
interface explicitly because two mistakes are otherwise easy: treating a
separated-angle light-ray density as a detector distribution at the endpoint,
and forgetting the geometric pushforward from two detector directions to one
angular variable.

Let
\[
  \rho=1-\mathbf n_1\cdot\mathbf n_2,\qquad 0<\rho<2,
\]
and let \(d_\perp=D-2\) be the dimension of the detector sphere.  For a fixed
second direction and a smooth test function \(F\) of the first direction,
geodesic polar coordinates give the exact pushforward identity
\begin{equation}
  \begin{aligned}
    &\int_{S^{d_\perp}}\dd\Omega_1\,
    \delta\!\left(\rho-(1-\mathbf n_1\cdot\mathbf n_2)\right)
    F(\mathbf n_1)\\
    &\quad =
    \mathcal J_{d_\perp}(\rho)
    \int_{S^{d_\perp-1}}\dd\omega\,
    F(\mathbf n(\rho,\omega)),\\
    &\mathcal J_{d_\perp}(\rho)
    =
    (2\rho-\rho^2)^{(d_\perp-2)/2}.
  \end{aligned}
  \label{eq:measured-small-angle-eec-pushforward}
\end{equation}
In four spacetime dimensions \(d_\perp=2\), so
\(\mathcal J_{2}=1\).  Thus the familiar \(1/\rho\) small-angle EEC
singularity is not produced by the area of a small detector annulus; it is a
short-distance light-ray or collinear singularity.  In other dimensions, or
in a detector patch with a nontrivial angular coordinate, the Jacobian is a
load-bearing part of the prediction.

Fix an endpoint interval \(0\le\rho\le\rho_0\), a finite retained
light-ray label set \(L\), and a smooth endpoint detector test \(\varphi\).
A small-angle contribution to the measured EEC is a distributional functional
of the form
\begin{equation}
  \left\langle
    O_{{\rm small},L}^{\rm pred},
    \varphi
  \right\rangle
  =
  \sum_{A\in L}
  \left\langle
    \mathcal J_{d_\perp} C_A^{\rm end}(\rho;Q,\mu,\nu),
    \varphi(\rho)
  \right\rangle
  \langle X|\mathbb O_A(\mu,\nu)|X\rangle
  +
  K_{+,L}(\varphi;X)
  +
  R_{{\rm small},L}(\varphi;X).
  \label{eq:measured-small-angle-eec-chart}
\end{equation}
Here \(C_A^{\rm end}\) is not an ordinary function at \(\rho=0\): it includes
the separated-angle density, endpoint plus-subtractions, and contact
coordinates such as the ones in
\eqref{eq:qcd-eec-renormalized-endpoint-coefficient}.  The term
\(K_{+,L}\) records explicit coincident-detector contact and endpoint-atom
coordinates after the detector convention has been chosen.  The remainder
\(R_{{\rm small},L}\) contains omitted light-ray labels, mixing curvature or
path-ordering defects, endpoint matching, and any regulator or rapidity
residuals not represented by the retained chart.

The endpoint-resolution identity
\eqref{eq:qcd-eec-plus-resolution-shift} is the operational sign check.  If
an annulus \(\rho_a<\rho<\rho_b\) is moved from the separated-angle bulk into
the endpoint plus-chart, the ordinary annulus integral is added and the
\(\delta(\rho)\) contact coordinate shifts by the compensating logarithmic
amount.  A finite detector bin touching \(\rho=0\) is therefore not predicted
by integrating only the displayed separated-angle density over the bin.  It is
predicted by pairing the full endpoint distribution with the detector test.

Combining \eqref{eq:finite-resolution-eec-observable-budget} with the
small-angle chart gives the local measured-bin requirement
\begin{equation}
  \left|
  \left\langle
    O_{\rm phys}-O_{{\rm small},L}^{\rm pred},
    \varphi
  \right\rangle
  \right|
  \le
  B_{\rm bin}
  +B_{\rm jac}
  +B_{\rm contact}
  +B_{\rm mix}
  +B_{\rm match}
  +B_{\rm np}
  +B_{\rm cont}.
  \label{eq:measured-small-angle-eec-budget}
\end{equation}
The new entries here are not decorative.  \(B_{\rm jac}\) records angular
pushforward and detector-coordinate errors; \(B_{\rm contact}\) records the
endpoint plus/contact convention; \(B_{\rm mix}\) records omitted light-ray
operators or non-flat retained transport; \(B_{\rm match}\) records matching
to the separated-angle and back-to-back charts; and \(B_{\rm np}\) records
hadronization, track, or fragmentation data.  Thus the physics prediction is a
statement about a measured detector functional.  A list of small-angle powers
or anomalous dimensions becomes such a prediction only after the pushforward,
endpoint contact transport, retained-operator residual, and finite-resolution
binning have been supplied in the same convention.
\end{controlledapproximation}

\begin{controlledapproximation}[Measured EEC global moment closure]
\label{ca:measured-eec-global-moment-closure}
The local endpoint chart is not the whole measured EEC.  Small-angle and
back-to-back factorizations can each be accurate in their own endpoint
coordinates while their union fails the global detector moment identities.  The
closure condition is elementary, but it is load-bearing because it ties the
endpoint atoms to measured energy and momentum rather than to a convenient
partonic normalization.

Use the angular variable
\(\zeta=\mathbf n_1\cdot\mathbf n_2\in[-1,1]\).  For an event with energy
fractions \(z_r=E_r/Q\), directions \(\mathbf n_r\), and selector weights
\(0\leq\omega_r\leq1\), define
\begin{equation}
  A_\omega
  =
  \left(\sum_r\omega_r z_r\right)^2,
  \qquad
  B_\omega
  =
  \left|\sum_r\omega_r z_r\mathbf n_r\right|^2 .
  \label{eq:measured-eec-selected-moments}
\end{equation}
For a full calorimeter in the center-of-momentum frame,
\(A_\omega=1\) and \(B_\omega=0\).  For tracks, flavor tags, fiducial regions,
or any other selected-energy observable, \(A_\omega\) and \(B_\omega\) are
measured inputs or ensemble averages; replacing them by the full-calorimeter
values changes the observable.

Let \(O_\omega^{\rm open}\) be the open-chart contribution on
\(-1<\zeta<1\), after the separated-angle, small-angle, and back-to-back
subtraction conventions have been chosen.  A contact-inclusive endpoint
assembly has the form
\begin{equation}
  O_\omega^{\rm pred}
  =
  O_\omega^{\rm open}
  +
  K_+^\omega\,\delta(1-\zeta)
  +
  K_-^\omega\,\delta(1+\zeta)
  +
  \mathcal R_\omega ,
  \label{eq:measured-eec-global-endpoint-decomposition}
\end{equation}
where \(K_+^\omega\) is the coincident-detector atom and \(K_-^\omega\) is the
back-to-back atom.  Define the open moments
\begin{equation}
  I_0^\omega=\left\langle O_\omega^{\rm open},1\right\rangle,
  \qquad
  I_1^\omega=\left\langle O_\omega^{\rm open},\zeta\right\rangle .
  \label{eq:measured-eec-open-moments}
\end{equation}
Moment closure requires
\begin{equation}
  I_0^\omega+K_+^\omega+K_-^\omega=A_\omega,
  \qquad
  I_1^\omega+K_+^\omega-K_-^\omega=B_\omega .
  \label{eq:measured-eec-moment-closure-equations}
\end{equation}
Thus the endpoint atoms are not free matching constants:
\begin{equation}
  K_+^\omega
  =
  \frac12\left(A_\omega-I_0^\omega+B_\omega-I_1^\omega\right),
  \qquad
  K_-^\omega
  =
  \frac12\left(A_\omega-I_0^\omega-B_\omega+I_1^\omega\right).
  \label{eq:measured-eec-endpoint-atom-solution}
\end{equation}

Equivalently, if a proposed short-distance assembly supplies
\(\widehat I_a^\omega\) and \(\widehat K_\pm^\omega\), its global moment
defects are
\begin{equation}
  R_0^\omega
  =
  A_\omega-\widehat I_0^\omega-\widehat K_+^\omega-\widehat K_-^\omega,
  \qquad
  R_1^\omega
  =
  B_\omega-\widehat I_1^\omega-\widehat K_+^\omega+\widehat K_-^\omega .
  \label{eq:measured-eec-moment-defects}
\end{equation}
The unique endpoint-only correction that repairs both measured moments is
\begin{equation}
  \Delta K_+^\omega=\frac12(R_0^\omega+R_1^\omega),
  \qquad
  \Delta K_-^\omega=\frac12(R_0^\omega-R_1^\omega).
  \label{eq:measured-eec-moment-closure-correction}
\end{equation}
For a detector test \(\varphi\), this correction obeys the finite bound
\begin{equation}
  \left|
    \Delta K_+^\omega\varphi(1)
    +
    \Delta K_-^\omega\varphi(-1)
  \right|
  \leq
  \frac12|R_0^\omega+R_1^\omega|\,|\varphi(1)|
  +
  \frac12|R_0^\omega-R_1^\omega|\,|\varphi(-1)| .
  \label{eq:measured-eec-moment-closure-test-bound}
\end{equation}
If this correction is not imposed exactly, the residual proof-obligation map in
\eqref{eq:finite-resolution-eec-observable-budget} and
\eqref{eq:measured-small-angle-eec-budget} must include the corresponding
moment-closure term.  This is a global statement in \(\zeta\): the small-angle
light-ray OPE, the back-to-back recoil factorization, and the separated-angle
bulk are physical detector predictions only after their endpoint atoms glue to
the same measured \(A_\omega\) and \(B_\omega\).
\end{controlledapproximation}

\begin{controlledapproximation}[Measured EEC response and covariance contract]
\label{ca:measured-eec-response-covariance-contract}
Moment closure is still a one-event statement.  A collider measurement reports
finite detector bins and their ensemble covariance after a detector response,
so the physical prediction needs one more layer before it is comparable with
data.  Fix detector tests \(\varphi_\alpha\) on \([-1,1]\).  For an event
\(e\), define the contact-inclusive vector
\begin{equation}
  X_\alpha(e)
  =
  \left\langle O_{\omega,e}^{\rm open},\varphi_\alpha\right\rangle
  +
  K^\omega_+(e)\varphi_\alpha(1)
  +
  K^\omega_-(e)\varphi_\alpha(-1) .
  \label{eq:measured-eec-event-vector}
\end{equation}
The endpoint atoms in \eqref{eq:measured-eec-event-vector} are eventwise
coordinates.  Solving \eqref{eq:measured-eec-moment-closure-correction} only
after ensemble averaging repairs the first two mean moments, but it does not
repair bin covariances whenever the selected energy, recoil, or endpoint
defect fluctuates across events.
For a response matrix built from calorimeter cells rather than from a
one-variable \(\zeta\) histogram, \(X_\alpha(e)\) is formed before the
opening-angle pushforward: the same-hadron part is the diagonal detector-cell
measure, and the scalar \(K_+^\omega(e)\delta(1-\zeta)\) is its final
projection.  This ordering fixes the covariance of same-cell products and
keeps a scalar endpoint repair from being reused across incompatible detector
coordinates.

Let \(\mathbb E\) denote the physical event ensemble.  Write
\begin{equation}
  \bar X_\alpha=\mathbb E[X_\alpha],
  \qquad
  \Sigma^X_{\alpha\beta}
  =
  \mathbb E[(X_\alpha-\bar X_\alpha)(X_\beta-\bar X_\beta)] .
  \label{eq:measured-eec-event-covariance}
\end{equation}
If a finite detector response maps the ideal vector to observed bins by
\begin{equation}
  Y_a
  =
  \sum_\alpha R_{a\alpha}X_\alpha+\epsilon_a,
  \qquad
  \mathbb E[\epsilon_a|X]=0,\qquad
  \mathbb E[\epsilon_a\epsilon_b|X]=N_{ab}(X),
  \label{eq:measured-eec-linear-response-model}
\end{equation}
then the measured mean and covariance are
\begin{equation}
  \bar Y=R\bar X,
  \qquad
  \Sigma^Y=R\Sigma^X R^{T}+\mathbb E[N(X)] .
  \label{eq:measured-eec-response-covariance}
\end{equation}
The covariance layer is part of the observable datum.  The response matrix,
detector-noise coordinate, selected-energy normalization, and endpoint
convention must be transported together.  A theory curve for binned EEC data
should therefore state a paired mean-and-covariance contract: for all finite
bin covectors \(v\),
\begin{equation}
  \begin{aligned}
  \left|v\cdot(\bar Y_{\rm phys}-R\bar X_{\rm pred})\right|
  &\le B_{\rm mean}(v),\\
  \left|
  v^T\!
  \left(
    \Sigma^Y_{\rm phys}
    -R\Sigma^{X}_{\rm pred}R^T
    -\overline N
  \right)v
  \right|
  &\le B_{\rm cov}(v).
  \end{aligned}
  \label{eq:measured-eec-mean-covariance-contract}
\end{equation}
Here \(B_{\rm mean}\) contains the same binning, endpoint, perturbative,
Glauber, nonperturbative, track, and continuum terms as
\eqref{eq:finite-resolution-eec-observable-budget}, while \(B_{\rm cov}\)
contains their covariance counterparts and the uncertainty in the response
calibration \(\overline N=\mathbb E[N(X)]\).

The eventwise endpoint repair has a finite covariance identity.  Let
\(\widehat X(e)\) be a proposed open-plus-endpoint vector and let
\[
  C_\alpha(e)
  =
  \Delta K^\omega_+(e)\varphi_\alpha(1)
  +
  \Delta K^\omega_-(e)\varphi_\alpha(-1)
\]
be the moment-closure correction from
\eqref{eq:measured-eec-moment-closure-correction}.  Then
\begin{equation}
  \Sigma^{\widehat X+C}
  =
  \Sigma^{\widehat X}
  +\Sigma^{\widehat X,C}
  +\Sigma^{C,\widehat X}
  +\Sigma^C .
  \label{eq:measured-eec-endpoint-covariance-repair}
\end{equation}
Discarding the three correction terms in
\eqref{eq:measured-eec-endpoint-covariance-repair} is the covariance analogue
of using \(A_\omega=1,B_\omega=0\) for a selected-energy observable.  It may
leave the mean curve visually plausible while mis-normalizing the uncertainty,
bin-to-bin correlations, and any correlated \(\chi^2\) comparison.  The
observable-level claim is therefore not ``the light-ray chart has the right
endpoint powers.''  It is that the eventwise chart, endpoint atoms,
nonperturbative lift, and detector response together satisfy
\eqref{eq:measured-eec-mean-covariance-contract}.
\end{controlledapproximation}

\section{Shape Functions and Track Functions}

When a measured event shape is of order \(\Lambda_{\rm QCD}/Q\), the soft
function in \eqref{eq:jet-event-shape-factorization} contains
nonperturbative matrix elements.  The endpoint statement has two logically
separate pieces.  First one must define a soft operator coordinate.  Then one
may state a factorized approximation that convolves this coordinate with a
short-distance coefficient in a specified endpoint topology.

\begin{definition}[Leading shape-function operator definition]
\label{def:leading-shape-function-operator-datum}
Fix a two-jet SCET operator specification as in
Definition~\ref{def:scet-two-jet-operator-datum}, an event-shape soft
measurement coordinate \(\widehat\ell_e\), and a renormalization scheme
\(\mathcal R_\mu\) for the cusp and endpoint singularities of the Wilson-line
operator.  A leading shape-function operator definition consists of:
\begin{enumerate}
  \item outgoing soft Wilson lines \(Y_n,Y_{\bar n}\) in the representations
  carried by the two energetic charges;
  \item for every test function \(h\) on \([0,\infty)\), a renormalized
  composite insertion \(h_{\mathcal R_\mu}(\widehat\ell_e)\);
  \item a normalization factor \(N_R\), equal to the dimension of the color
  representation in the fundamental two-jet case, together with the convention
  that the identity insertion has unit matrix element.
\end{enumerate}
The shape function \(F_e(\ell,\mu;\mathcal R)\) is the distribution on
\([0,\infty)\) defined by
\begin{equation}
  \int_0^\infty \dd\ell\, h(\ell)F_e(\ell,\mu;\mathcal R)
  =
  \frac1{N_R}
  \operatorname{tr}
  \bra{\vac}
  \overline{\operatorname T}\!\left(Y_{\bar n}^{\dagger}Y_n\right)
  h_{\mathcal R_\mu}(\widehat\ell_e)
  \operatorname T\!\left(Y_n^{\dagger}Y_{\bar n}\right)
  \ket{\vac}_{\mathcal R_\mu}.
  \label{eq:shape-function-test-pairing}
\end{equation}
At finite regulator with a positive self-adjoint measurement operator
\(\widehat\ell_e\), this pairing is the spectral measure of
\(\widehat\ell_e\) in the Wilson-line state.  After renormalization it is a
scheme-dependent distributional coordinate.  The normalization and first
moment are
\begin{equation}
  \int_0^\infty\dd\ell\,F_e(\ell,\mu;\mathcal R)=1,
  \qquad
  \Omega_1^e(\mu;\mathcal R)
  =
  \int_0^\infty\dd\ell\,\ell\,F_e(\ell,\mu;\mathcal R),
  \label{eq:shape-function-normalization-moment}
\end{equation}
whenever the displayed moments exist in the chosen scheme.
\end{definition}

The QFT object is the operator pairing
\eqref{eq:shape-function-test-pairing}.  In a phenomenological analysis one
often chooses a finite-dimensional positive family for \(F_e\) and fits its
parameters to data.  That fitted family is an additional statistical model
for the matrix element \eqref{eq:shape-function-test-pairing}.
The lattice route to the same nonperturbative information must start from a
finite gauge-field regulator as in
Chapter~\ref{ch:lattice-yang-mills-nonperturbative-definition} and a
continuum reconstruction or matching statement as in
Chapter~\ref{ch:constructive-lattice-continuum-local-qft}.  Because
\eqref{eq:shape-function-test-pairing} contains lightlike Wilson-line
geometry, a Euclidean calculation needs a separately specified quasi-,
ratio-, or reconstruction observable and a matching theorem connecting it to
the lightlike shape coordinate.

\begin{controlledapproximation}[Leading endpoint shape-function convolution]
\label{ca:leading-endpoint-shape-function-convolution}
Let \(e\) be a dimensionless two-jet event shape and let \(Q\) be the hard
scale.  A leading shape-function description on an endpoint window
\(e\sim \Lambda_{\rm QCD}/Q\) consists of:
\begin{enumerate}
  \item the operator definition of
  Definition~\ref{def:leading-shape-function-operator-datum};
  \item a perturbative coefficient distribution
  \(\dd\sigma_{\rm pert}/\dd e\) in the same subtraction scheme;
  \item a Banach space \(\mathcal B_{\rm end}\) of test functions supported
  in the endpoint window;
  \item a remainder functional \(R_Q\) on \(\mathcal B_{\rm end}\) with a
  stated bound, for example
  \(\lVert R_Q\rVert_{\mathcal B_{\rm end}^\ast}
  \leq C Q^{-2}\) in a single-soft-coordinate situation.
\end{enumerate}
The approximation means that, for every \(f\in\mathcal B_{\rm end}\),
\begin{equation}
  \left\langle \frac{\dd\sigma}{\dd e},f\right\rangle
  =
  \int_0^\infty\dd\ell\,F_e(\ell,\mu;\mathcal R)
  \int\dd e'\,
  \frac{\dd\sigma_{\rm pert}}{\dd e'}(e',Q,\mu;\mathcal R)\,
  f\!\left(e'+\frac{\ell}{Q}\right)
  +R_Q(f).
  \label{eq:shape-function-distributional-convolution}
\end{equation}
Equivalently, when the distributions admit densities in the chosen chart,
\begin{equation}
  \frac{\dd\sigma}{\dd e}
  =
  \int_0^\infty\dd\ell\,
  F_e(\ell,\mu;\mathcal R)
  \frac{\dd\sigma_{\rm pert}}{\dd e}
  \left(e-\frac{\ell}{Q},Q,\mu;\mathcal R\right)
  +R_Q(e).
  \label{eq:shape-function-density-convolution}
\end{equation}
\end{controlledapproximation}

The elementary consequences of the convolution are useful diagnostics for
normalization and scheme bookkeeping.  Taking \(f(e)=1\) in
\eqref{eq:shape-function-distributional-convolution} and using
\eqref{eq:shape-function-normalization-moment} gives equality of the total
weighted cross section up to \(R_Q(1)\).  Taking \(f(e)=e\) gives
\begin{equation}
  \int\dd e\,e\,\frac{\dd\sigma}{\dd e}
  =
  \int\dd e\,e\,\frac{\dd\sigma_{\rm pert}}{\dd e}
  +
  \frac{\Omega_1^e(\mu;\mathcal R)}{Q}
  \int\dd e\,\frac{\dd\sigma_{\rm pert}}{\dd e}
  +R_Q(e).
  \label{eq:shape-function-first-moment-shift}
\end{equation}
Thus the first nonperturbative moment translates the endpoint event shape by
\(\Omega_1^e/Q\) in this leading single-coordinate approximation.  A finite
change of subtraction convention may move a constant amount \(\delta\ell\)
between the perturbative coefficient and the shape coordinate:
\begin{equation}
  F'_e(\ell)=F_e(\ell-\delta\ell),
  \qquad
  \frac{\dd\sigma'_{\rm pert}}{\dd e}(e)
  =
  \frac{\dd\sigma_{\rm pert}}{\dd e}
  \left(e+\frac{\delta\ell}{Q}\right).
  \label{eq:shape-function-scheme-translation}
\end{equation}
The convolution is invariant under the paired transformation
\eqref{eq:shape-function-scheme-translation}, modulo boundary and truncation
terms.  Hence a numerical value quoted for \(\Omega_1^e\) is meaningful only
with its event-shape definition, Wilson-line orientation, regulator, and
subtraction scheme.

\begin{controlledapproximation}[Shape-function moment expansion for smeared endpoint tests]
\label{ca:shape-function-smeared-moment-expansion}
The first-moment shift is the first term in a smeared-test expansion, not a
replacement for the full nonperturbative distribution.  Assume that the shape
coordinate in Definition~\ref{def:leading-shape-function-operator-datum} has a
finite second moment
\[
  \Omega_2^e(\mu;\mathcal R)
  =
  \int_0^\infty\dd\ell\,\ell^2F_e(\ell,\mu;\mathcal R)
\]
and let \(f\in C^2\) be supported in the declared endpoint window.  For each
perturbative endpoint value \(e'\),
\[
  f\!\left(e'+\frac{\ell}{Q}\right)
  =
  f(e')
  +
  \frac{\ell}{Q}f'(e')
  +
  \int_0^{\ell/Q}\!
    \left(\frac{\ell}{Q}-t\right)f''(e'+t)\,\dd t .
\]
Substitution into
\eqref{eq:shape-function-distributional-convolution} gives
\begin{equation}
\begin{aligned}
  \left\langle \frac{\dd\sigma}{\dd e},f\right\rangle
  &=
  \left\langle
    \frac{\dd\sigma_{\rm pert}}{\dd e},
    f+\frac{\Omega_1^e}{Q}f'
  \right\rangle
  +
  \mathcal E_{2,Q}^e[f]
  +
  R_Q(f),
\\
  |\mathcal E_{2,Q}^e[f]|
  &\le
  \frac{\Omega_2^e}{2Q^2}
  \left\|\frac{\dd\sigma_{\rm pert}}{\dd e}\right\|_{\rm TV}
  \|f''\|_\infty ,
\end{aligned}
  \label{eq:shape-function-smeared-moment-expansion}
\end{equation}
where \(\|\dd\sigma_{\rm pert}/\dd e\|_{\rm TV}\) denotes the total variation
of the perturbative endpoint distribution in the chosen chart.  For a
positive normalized perturbative measure this factor is one.

Equation~\eqref{eq:shape-function-smeared-moment-expansion} is the practical
content of the sentence that a leading shape function shifts an event shape:
linear tests see only \(\Omega_1^e/Q\), while curved endpoint tests also see
\(\Omega_2^e/Q^2\) and the higher moments hidden in the residual.  Two fitted
families with the same normalization and first moment can therefore agree on
the total rate and the endpoint mean but disagree on a quadratic or localized
smearing.  If a retained moment model \(\widehat F_e^{(m)}\) replaces the
full Wilson-line distribution in a fit, the missing shape direction enters
the measured test family \(\mathcal F\subset\mathcal B_{\rm end}\) through
\begin{equation}
  B_{\rm shape}^{(m)}(\mathcal F)
  =
  \sup_{\substack{f\in\mathcal F\\ \|f\|_{\mathcal F}\le 1}}
  \left|
  \int_0^\infty \dd\ell\,
  \bigl(F_e(\ell)-\widehat F_e^{(m)}(\ell)\bigr)
  \int\dd e'\,
  \frac{\dd\sigma_{\rm pert}}{\dd e'}(e')\,
  f\!\left(e'+\frac{\ell}{Q}\right)
  \right| .
  \label{eq:shape-function-moment-truncation-budget}
\end{equation}
For polynomial tests of degree at most \(m\), matched moments can make
\(B_{\rm shape}^{(m)}\) vanish.  For finite endpoint bins, tail-sensitive
tests, or response windows with sharp edges, the same matched moments leave a
real shape-truncation budget.  A collider comparison that fits only
\(\Omega_1^e\), or only a few moments, therefore carries
\eqref{eq:shape-function-moment-truncation-budget} as an observable-level
uncertainty on the actual binned test family.  Conversely, using an
\(\Omega_1^e\) fitted in a different
subtraction convention changes even the linear endpoint mean unless the
paired translation
\eqref{eq:shape-function-scheme-translation} is applied.  The fit parameters
are coordinates for the Wilson-line matrix element
\eqref{eq:shape-function-test-pairing}, not new universal constants of QCD.
\end{controlledapproximation}

Track-based observables measure only charged hadrons.  They are generally not
IRC safe as ordinary partonic measurements, because a collinear splitting can
change the charged energy fraction observed by the detector.  The replacement
input is a track function.  For a parton species \(i\), let
\(T_i(x,\mu)\) be the renormalized track coordinate, in a specified operator
scheme, for the fraction \(x\in[0,1]\) of the parent energy carried by charged
tracks.  At a reference nonperturbative scale this coordinate is naturally a
probability measure.  After renormalization it should be viewed as a
scheme-dependent distributional coordinate whose normalization is fixed by the
operator definition; when it has a density we write the same symbol for that
density.  Perturbative coefficients are then convolved with \(T_i\) rather
than evaluated with a purely partonic charged-energy measurement.

\begin{definition}[Track-function operator definition]
\label{def:track-function-operator-datum}
Fix a lightlike collinear direction \(n\), the conjugate vector \(\bar n\)
with \(n\cdot\bar n=2\), and the SCET operator conventions of
Definition~\ref{def:scet-two-jet-operator-datum}.  A track-function operator
definition for a parton species \(i\) consists of:
\begin{enumerate}
  \item a gauge-invariant collinear source \(\mathcal O_i\), for example
  \(\chi_n\) for a quark species or
  \(\mathcal B_{n\perp}^{\mu}\) for a gluon species, together with the spin,
  color, and polarization averaging convention used to normalize the source;
  \item a detector weight \(w_{\rm tr}\) on stable color-singlet final-state
  particle species, with \(w_{\rm tr}=1\) for charged tracks and
  \(w_{\rm tr}=0\) for neutral particles in the idealized charged-track
  observable;
  \item a positive charged collinear momentum operator
  \(\bar n\cdot\widehat P_{\rm tr}\) obtained by weighting the stable
  final-state momentum operator with \(w_{\rm tr}\), and the total collinear
  momentum operator \(\bar n\cdot\widehat P_{\rm col}\);
  \item the bounded measurement operator
  \[
    \widehat x_{\rm tr}
    =
    \frac{\bar n\cdot\widehat P_{\rm tr}}
         {\bar n\cdot\widehat P_{\rm col}}
  \]
  on the positive-energy sector created by \(\mathcal O_i\), where
  \(0\leq \widehat x_{\rm tr}\leq1\);
  \item a regulator and subtraction scheme \(\mathcal R_\mu\) for the
  light-ray collinear matrix element and for the endpoint singularities of
  the measurement insertion.
\end{enumerate}
For every test function \(h\) on \([0,1]\), the track function is the
renormalized distribution defined by
\begin{equation}
  \int_0^1\dd x\,h(x)T_i(x,\mu;\mathcal R)
  =
  \frac{1}{\mathcal N_i(\mu;\mathcal R)}
  \bra{\vac}
  \overline{\mathcal O}_i(0)\,
  h_{\mathcal R_\mu}(\widehat x_{\rm tr})\,
  \mathcal O_i(0)
  \ket{\vac}_{\mathcal R_\mu}.
  \label{eq:track-function-test-pairing}
\end{equation}
The normalization factor \(\mathcal N_i\) is fixed by the requirement that
the pairing with \(h=1\) equals \(1\).  At finite regulator this is the
spectral distribution of a bounded charged-energy fraction in the inclusive
collinear state created by the gauge-invariant source.  In continuum QCD it
is a renormalized distributional matrix element; its value is meaningful only
after the source, detector weight, regulator, and subtraction scheme have
all been specified.
\end{definition}

This definition is the nonperturbative QCD object associated with the chosen
collinear source.  The parton label \(i\) labels that source in the
factorized approximation.  Gauge invariance enters through the Wilson-line
building blocks \(\chi_n\) and \(\mathcal B_{n\perp}\), while the
charged-track measurement is a color-singlet final-state measurement.  The
renormalization of \eqref{eq:track-function-test-pairing} is separate from
the renormalization of the hard partonic coefficient into which the
track function is inserted.

\begin{definition}[Paired track-function evolution structure]
\label{def:paired-track-function-evolution}
Fix a renormalization scale \(\mu\).  A finite-kernel track-function evolution
structure consists of:
\begin{itemize}
  \item a finite set \(\mathcal I\) of parton species;
  \item for every ordered splitting \(i\to jk\), an integrable kernel
  \(K_{i\to jk}(z,\mu)\) on \(0<z<1\);
  \item the paired loss rate
  \[
    \Gamma_i(\mu)
    =
    \sum_{j,k\in\mathcal I}\int_0^1\dd z\,K_{i\to jk}(z,\mu).
  \]
\end{itemize}
The associated evolution equation for track coordinates is
\begin{align}
  \mu\frac{\dd}{\dd\mu}T_i(x,\mu)
  &=
  \sum_{j,k\in\mathcal I}
  \int_0^1\dd z
  \int_0^1\dd x_1\,\dd x_2\,
  K_{i\to jk}(z,\mu)\,
  T_j(x_1,\mu)T_k(x_2,\mu)
  \delta\!\left(x-zx_1-(1-z)x_2\right)
  \nonumber\\
  &\hspace{2.5em}
  -\Gamma_i(\mu)T_i(x,\mu).
  \label{eq:paired-track-function-evolution}
\end{align}
In perturbative QCD the kernels are renormalized splitting distributions
rather than honest integrable functions before smearing; the finite-kernel
definition records the algebraic real--virtual pairing that the
renormalized distributional equation must preserve.
\end{definition}

\paragraph{Normalization and first moment of track-function evolution.}
Suppose the track functions in
Definition~\ref{def:paired-track-function-evolution} have unit normalization,
\[
  \int_0^1\dd x\,T_i(x,\mu)=1 .
\]
The paired evolution
\eqref{eq:paired-track-function-evolution} preserves this normalization.
If
\[
  m_i(\mu)=\int_0^1\dd x\,x\,T_i(x,\mu)
\]
is the charged-energy first moment, then
\begin{equation}
  \mu\frac{\dd}{\dd\mu}m_i(\mu)
  =
  \sum_{j,k}\int_0^1\dd z\,K_{i\to jk}(z,\mu)
  \left[z\,m_j(\mu)+(1-z)m_k(\mu)-m_i(\mu)\right].
  \label{eq:track-function-first-moment-rg}
\end{equation}
Integrating \eqref{eq:paired-track-function-evolution} over \(x\) and using
\[
  \int_0^1\dd x\,
  \delta\!\left(x-zx_1-(1-z)x_2\right)
  =
  1,
\]
because \(0\leq zx_1+(1-z)x_2\leq1\), gives
\[
  \mu\frac{\dd}{\dd\mu}\int_0^1\dd x\,T_i(x,\mu)
  =
  \sum_{j,k}\int_0^1\dd z\,K_{i\to jk}(z,\mu)
  \left[
    \left(\int T_j\right)\left(\int T_k\right)
    -\int T_i
  \right].
\]
If all normalizations are \(1\), the bracket vanishes for every splitting
channel.  Hence the normalization is preserved.

Multiplying by \(x\) before integrating gives
\[
  \int_0^1\dd x\,x\,
  \delta\!\left(x-zx_1-(1-z)x_2\right)
  =
  zx_1+(1-z)x_2.
\]
The gain term therefore contributes
\[
  \sum_{j,k}\int_0^1\dd z\,K_{i\to jk}(z,\mu)
  \left[
    z\,m_j(\mu)\int T_k
    +(1-z)\,m_k(\mu)\int T_j
  \right].
\]
Using unit normalization and subtracting the loss term
\(\Gamma_i m_i\) gives \eqref{eq:track-function-first-moment-rg}.

The same finite-kernel equation controls the whole moment tower.  Define
\[
  m_i^{(r)}(\mu)=\int_0^1\dd x\,x^rT_i(x,\mu),
  \qquad
  m_i^{(0)}=1,
\]
and, when the exponential test function is admissible in the chosen
distributional scheme,
\[
  M_i(\nu,\mu)=\int_0^1\dd x\,e^{\nu x}T_i(x,\mu).
\]
Pairing \eqref{eq:paired-track-function-evolution} with \(e^{\nu x}\) gives
\begin{equation}
  \mu\frac{\dd}{\dd\mu}M_i(\nu,\mu)
  =
  \sum_{j,k}\int_0^1\dd z\,K_{i\to jk}(z,\mu)
  \left[
    M_j(z\nu,\mu)M_k((1-z)\nu,\mu)-M_i(\nu,\mu)
  \right].
  \label{eq:track-function-generating-rg}
\end{equation}
Expanding both sides in powers of \(\nu\) yields, for every \(r\geq0\),
\begin{equation}
  \mu\frac{\dd}{\dd\mu}m_i^{(r)}
  =
  \sum_{j,k}\int_0^1\dd z\,K_{i\to jk}(z,\mu)
  \left[
    \sum_{a=0}^{r}
    \binom{r}{a}
    z^a(1-z)^{r-a}
    m_j^{(a)}m_k^{(r-a)}
    -m_i^{(r)}
  \right].
  \label{eq:track-function-moment-tower-rg}
\end{equation}
The \(r=0\) member is normalization preservation and the \(r=1\) member is
\eqref{eq:track-function-first-moment-rg}.  For \(r\geq2\) the evolution of a
single moment depends on lower and equal moments of daughter species.  This
is why track-based jet observables require a genuine distributional
nonperturbative input, or an equivalently complete moment sequence in a
specified scheme; a single average charged fraction is not stable under
collinear renormalization.

Track functions are therefore nonperturbative inputs with perturbative
evolution, not substitutes for IRC safety.  The point is sharper than a
terminological distinction: a track observable is calculable only after the
observable has been lifted from a partonic measurement to a convolution with
the nonperturbative track coordinates and the real--virtual pairing in
Definition~\ref{def:paired-track-function-evolution} has been specified.

\begin{definition}[Finite track lift of an energy-measure observable]
\label{def:finite-track-observable-lift}
Fix a partonic short-distance configuration with labeled partons
\[
  a=1,\ldots,n,\qquad (i_a,E_a,y_a),
\]
where \(i_a\) is the parton species, \(E_a>0\) is the energy coordinate, and
\(y_a\) is the angular coordinate in the detector patch.  A finite
track-function lift assigns to each label \(a\) an independent random variable
\(X_a\in[0,1]\) with law \(T_{i_a}\) and replaces the calorimetric energy
measure by the random charged-track energy measure
\[
  \mu_{\rm tr}
  =
  \sum_{a=1}^n X_a E_a\,\delta_{y_a}.
\]
For a smeared \(k\)-point energy polynomial with continuous kernel \(f\),
\[
  \mathcal E_f(\mu)
  =
  \int f(y_1,\ldots,y_k)\,\mu(dy_1)\cdots\mu(dy_k),
\]
the lifted partonic coefficient is the conditional expectation
\[
  \mathbb E_T\,\mathcal E_f(\mu_{\rm tr}).
\]
This finite definition is the algebraic core of a track-based factorization
formula.  A QCD prediction additionally requires the operator definition of
the track functions, their subtraction scheme and RG evolution, the
short-distance coefficient into which this lift is inserted, and a remainder
estimate for the leading-power track-factorization approximation.
\end{definition}

The finite lift shows why track observables cannot be reduced to replacing
each energy by the average charged fraction.  Let
\[
  m_i^{(r)}=\int_0^1 x^r\,T_i(x)\,\dd x
\]
be the \(r\)-th track moment whenever this moment is defined in the chosen
scheme.  For the two-point polynomial one obtains
\begin{align}
  \mathbb E_T\,\mathcal E_f(\mu_{\rm tr})
  &=
  \sum_{a\neq b}
  E_aE_b\,f(y_a,y_b)\,
  m_{i_a}^{(1)}m_{i_b}^{(1)}
  \nonumber\\
  &\hspace{2em}
  +
  \sum_a
  E_a^2\,f(y_a,y_a)\,
  m_{i_a}^{(2)} .
  \label{eq:track-lift-two-point-moment-formula}
\end{align}
The second line is a diagonal term in the energy-measure polynomial.  It is
controlled by the second track moment, not by the square of the first moment.
For a general \(k\)-point energy polynomial the exact finite formula is the
following.  Let
\[
  \alpha:\{1,\ldots,k\}\longrightarrow\{1,\ldots,n\}
\]
record which short-distance parton supplies the energy in each detector slot.
Independence of the finite track variables gives
\begin{equation}
  \mathbb E_T\,\mathcal E_f(\mu_{\rm tr})
  =
  \sum_{\alpha:\{1,\ldots,k\}\to\{1,\ldots,n\}}
  \left(\prod_{r=1}^{k}E_{\alpha(r)}\right)
  f\!\left(y_{\alpha(1)},\ldots,y_{\alpha(k)}\right)
  \prod_{a\in\operatorname{im}\alpha}
  m_{i_a}^{(|\alpha^{-1}(a)|)} .
  \label{eq:track-lift-general-k-point-formula}
\end{equation}
Thus each block of detector slots fed by the same short-distance label uses
the corresponding higher track moment.  For a three-point kernel, writing
\(f_{abc}=f(y_a,y_b,y_c)\), the formula separates into
\begin{align}
  \mathbb E_T\,\mathcal E_f(\mu_{\rm tr})
  &=
  \sum_{\substack{a,b,c\\\text{distinct}}}
  E_aE_bE_c\,f_{abc}\,
  m_{i_a}^{(1)}m_{i_b}^{(1)}m_{i_c}^{(1)}
  \nonumber\\
  &\quad+
  \sum_{a\ne b}
  E_a^2E_b\,
  \bigl(f_{aab}+f_{aba}+f_{baa}\bigr)\,
  m_{i_a}^{(2)}m_{i_b}^{(1)}
  \nonumber\\
  &\quad+
  \sum_a
  E_a^3\,f_{aaa}\,m_{i_a}^{(3)} .
  \label{eq:track-lift-three-point-moment-formula}
\end{align}
The diagonal and partial-diagonal pieces are not optional refinements: they
are part of the definition of the lifted observable.  Omitting them, or
replacing every track variable by its first moment, changes the observable
whose short-distance coefficient is being computed.

\begin{remark}[Track-function truncation for measured tests]
\label{rem:track-function-measured-test-warning}
The moment formulae above are exact only for the displayed polynomial
energy-measure tests.  A finite collider comparison normally reports
histogram bins, trigger regions, ratios, or a detector-response vector.  These
are tests on the random charged-track measure
\[
  \mu_{\rm tr}(X)=\sum_a X_aE_a\delta_{y_a},
\]
so they depend on the finite-bin law of the charged fractions.  If the full
track coordinates \(T_i\) are replaced by a finite moment model
\(\widehat T_i^{(m)}\) matched through moments \(0,\ldots,m\), then the
missing track-distribution direction enters a declared measured test family
\(\mathcal F\) through
\begin{equation}
  B_{\rm tr}^{(m)}(\mathcal F\,|\,C)
  =
  \sup_{\substack{\Phi\in\mathcal F\\ \|\Phi\|_{\mathcal F}\le1}}
  \left|
    \mathbb E_{T}\,
    \Phi\!\left(\sum_a X_aE_a\delta_{y_a}\right)
    -
    \mathbb E_{\widehat T^{(m)}}\,
    \Phi\!\left(\sum_a \widehat X_aE_a\delta_{y_a}\right)
  \right| ,
  \label{eq:track-function-measured-test-budget}
\end{equation}
where \(C=\{(i_a,E_a,y_a)\}\) is the retained short-distance configuration
and the variables are independently drawn from the corresponding track
coordinates in the finite lift.  This display defines a measured-test
discrepancy; it is not a controlled approximation unless one also supplies a
metric-determining test class and a quantitative estimate of the moment
truncation in that topology.  For polynomial \(\Phi\) whose repeated-label
blocks have size at most \(m\), moment matching can make
\(B_{\rm tr}^{(m)}\) vanish.  For a charged-energy threshold, a finite
histogram bin, or a response window with sharp edges, the same low moments do
not determine the probability of the bin.

The covariance part of a track-based measurement carries the same warning.  If
\(X_\alpha(e)\) denotes the ideal finite test vector for event \(e\), replacing
\(T_i\) by \(\widehat T_i^{(m)}\) changes not only \(\mathbb E[X_\alpha]\) but
also
\[
  \Sigma^X_{\alpha\beta}
  =
  \mathbb E[(X_\alpha-\bar X_\alpha)(X_\beta-\bar X_\beta)] .
\]
Thus a track prediction compared with charged-track data must include, in the
same convention as \eqref{eq:measured-eec-mean-covariance-contract}, both a
mean residual \(B_{\rm tr}^{(m)}\) and the corresponding covariance residual
on the reported test vectors.  A fitted average charged fraction, or even a
finite list of matched moments, is a coordinate approximation to the
nonperturbative track distribution; it is not by itself a prediction for all
charged-track bins.
\end{remark}

The same bookkeeping explains the collinear composition law behind
\eqref{eq:paired-track-function-evolution}.  If a parent with energy \(E\)
is resolved into daughters with energy fractions \(z\) and \(1-z\), then the
track fraction assigned to the unresolved pair is
\[
  X=zX_1+(1-z)X_2 .
\]
Thus
\begin{align}
  \mathbb E[X]
  &=
  z\,m_j^{(1)}+(1-z)m_k^{(1)},\\
  \mathbb E[X^2]
  &=
  z^2m_j^{(2)}+(1-z)^2m_k^{(2)}
  +2z(1-z)m_j^{(1)}m_k^{(1)}.
  \label{eq:track-lift-collinear-second-moment}
\end{align}
A track-based observable is therefore not IRC safe as a bare partonic
function.  Rather, its collinear boundary is a renormalization problem for
the whole sequence of track moments, or equivalently for the track
distribution.  This is the precise meaning of using track functions in a
jet-substructure calculation.

\section{Parton Showers}

A parton shower is a stochastic implementation of soft and collinear
factorized evolution.  At leading soft-collinear order, a final-state
branching probability for a parton of type \(i\) is built from the Sudakov
factor
\begin{equation}
  \Delta_i(t_1,t_0)
  =
  \exp\left[
    -\int_{t_0}^{t_1}\frac{\dd t}{t}
    \sum_{i\to jk}
    \int \dd z\,
    \frac{\alpha_s(\kappa(t,z))}{2\pi}
    P_{i\to jk}(z)
  \right],
  \label{eq:parton-shower-sudakov}
\end{equation}
where \(t\) is the shower ordering variable, \(z\) is the splitting fraction,
\(P_{i\to jk}\) is an Altarelli--Parisi splitting kernel, and
\(\kappa(t,z)\) is the scale at which the running coupling is evaluated.
The probability for the next branching between \(t\) and \(t+\dd t\) is the
Sudakov no-branching probability multiplied by the differential splitting
kernel.

Different showers choose different evolution variables, recoil maps, momentum
conservation prescriptions, and color structures.  Angular-ordered showers
encode color coherence through emission angles.  Dipole and antenna showers
use color-connected pairs as emitters.  Many practical implementations use a
leading-color approximation in which the color algebra is replaced by a
planar color-flow structure; subleading-color corrections require additional
operator information.

\begin{controlledapproximation}[Shower accuracy]
A shower calculation has a stated logarithmic accuracy only for a specified
class of observables and a specified matching or merging scheme.  The basic
Markov shower reproduces leading soft-collinear logarithms for observables
compatible with its ordering variable.  NLO-matched prescriptions such as
MC@NLO or POWHEG replace the first hard emission and inclusive normalization
by exact NLO matrix-element information while retaining the shower resummation
for subsequent unresolved radiation.  Multijet merging combines samples with
different resolved parton multiplicities using a merging scale whose residual
dependence is part of the perturbative uncertainty estimate.
\end{controlledapproximation}

\section{Hadronization and Fragmentation Functions}

Hadronization is the nonperturbative formation of outgoing hadrons from the
high-energy QCD state produced by the short-distance interaction.  In the
exact theory it is simply time evolution in the physical Hilbert space of
QCD.  In phenomenological calculations it is represented by additional
nonperturbative inputs after the perturbative hard, jet, and soft evolution
has been organized.

For identified hadron production, the corresponding input is a fragmentation
function.  Let \(n^\mu\) be a lightlike vector and let
\(\bar n^\mu\) satisfy \(n\cdot\bar n=2\).  A quark fragmentation function may
be represented by a light-ray operator with Wilson lines:
\begin{equation}
  D_{h/q}(z,\mu)
  =
  \frac{1}{2zN_c}
  \sum_X
  \int\frac{\dd y^+}{2\pi}
  \ee^{-\ii P_h^- y^+/z}
  \operatorname{tr}
  \left[
    \frac{\gamma\cdot\bar n}{2}
    \langle0|W^\dagger(y^+)q(y^+)|hX\rangle
    \langle hX|\bar q(0)W(0)|0\rangle
  \right]_{\rm ren}.
  \label{eq:quark-fragmentation-function-operator}
\end{equation}
Here \(P_h^-\) is the large light-cone momentum of the observed hadron,
\(z=P_h^-/k^-\) is the momentum fraction relative to the parent partonic
coordinate \(k^-\), \(W\) is the appropriate lightlike Wilson line, and the
subscript indicates operator renormalization.  This is a nonperturbative
matrix-element definition, not a probability for a physical colored quark to
exist as an external state.  A regulator is part of the definition of the
light-ray composite, and lightlike Wilson lines may require endpoint, cusp, or
rapidity subtractions depending on the observable.  The scale dependence of
\(D_{h/q}\) is governed by DGLAP evolution, while its boundary value at a finite
scale is nonperturbative data.

Lund string and cluster hadronization models are phenomenological
parametrizations of this nonperturbative stage.  A string model uses color
flux-tube breaking with fitted parameters for string tension, flavor
production, and transverse momentum.  A cluster model first organizes the
partonic state into color-singlet clusters and then decays those clusters into
hadrons with fitted rules.  Their predictions therefore carry model and tuning
uncertainties in addition to perturbative uncertainties.

The lattice formulation of
Chapter~\ref{ch:lattice-yang-mills-nonperturbative-definition} is the
regulator-level nonperturbative definition of the gauge theory used in this
volume.  It computes Euclidean finite-volume correlation functions directly in
terms of hadronic states.  Real-time fragmentation functions and fully
exclusive hadronization histories involve lightlike or timelike information,
so their extraction from Euclidean lattice data requires additional
reconstruction, finite-volume, or large-momentum matching ideas.  This is a
technical obstruction in the observable, not a license to replace the
nonperturbative QCD input by a perturbative calculation.

\section{Classes of QCD Final-State Observables}

The useful separation is by the mathematical input required by the observable.

\begin{description}[style=nextline,leftmargin=1.5em]
\item[Total inclusive rates and IRC-safe event shapes.]
The required input is renormalized fixed-order amplitudes, KLN cancellation,
and resummation when hierarchical scales are measured.

\item[IRC-safe jet cross sections.]
The required input is an IRC-safe jet algorithm, fixed-order hard functions,
PDFs for incoming hadrons, and logarithmic resummation where jet vetoes or
small radii introduce large logarithms.

\item[Energy correlators.]
The required input is the stress-tensor energy-flow operators on the physical
Hilbert space, together with angular test kernels defining continuous
functionals of the calorimetric measure.  The density theorem
\ref{thm:energy-correlator-polynomial-density} explains why smeared
energy-correlator polynomials form a natural coordinate system for
finite-resolution calorimetric observables.  Partonic perturbation theory is
used only after the detector observable has been defined.

\item[Identified hadron spectra.]
The required input is fragmentation functions or fracture functions, their
operator renormalization, and nonperturbative boundary data.

\item[Fully exclusive hadronization patterns.]
The required input is the full QCD final-state distribution or a
phenomenological model with stated fit parameters and model uncertainty.
\end{description}

This classification keeps the physics claims separated.  Perturbative QCD
computes short-distance coefficients and IRC-safe partonic representatives.
Fragmentation functions, PDFs, shape functions, and hadronization models are
additional nonperturbative data or controlled approximations with their own
definitions and uncertainties.  Energy correlators are especially valuable
because they are nonperturbatively defined by the stress tensor and are
insensitive, after smearing, to soft emissions and collinear recombination in
the precise sense used in
Controlled Approximation~\ref{capprox:irc-safety-fixed-order-finiteness}.
