\chapter{Bound States from Exchange Amplitudes}
\label{ch:volume-ii-bound-states-exchange}
\ConventionsLorentzian


\section{Exchange Model}

Consider two real scalar fields in four spacetime dimensions with Lorentzian
Lagrangian density
\[
  \mathcal L
  =
  -\frac12\partial_\mu\phi_1\partial^\mu\phi_1
  -\frac12m_1^2\phi_1^2
  -\frac12\partial_\mu\phi_2\partial^\mu\phi_2
  -\frac12m_2^2\phi_2^2
  -\frac12g\phi_1^2\phi_2 .
\]
The masses \(m_1,m_2\) are positive, and \(g\) is a real coupling. Use the
same mostly-plus convention as before. The scalar propagators and cubic
vertex are
\[
  \phi_i:\quad
  \frac{-\ii}{k^2+m_i^2-\ii0},
  \qquad
	  \phi_1\phi_1\phi_2:\quad -\ii g .
\]
This Lagrangian is used as a perturbative exchange model in the sense of
Section~\ref{sec:volume-ii-scalar-model-status}.  The pole criterion derived
below is a statement about any exact massive local theory whose amplitude has
the displayed analytic structure; the polynomial scalar density by itself is
only the finite-cutoff or effective description from which the tree kernel is
computed.

The connected \(\phi_1\phi_1\to\phi_1\phi_1\) tree amplitude receives
\(s\)-, \(t\)-, and \(u\)-channel exchange contributions.  The three routings
and the solid/dashed line convention are fixed in
Figure~\ref{fig:exchange-model-stu-tree-routings}.
\begin{figure}[!ht]
\centering
\begin{tikzpicture}[baseline=-0.5ex,scale=0.86,
  one/.style={line width=0.9pt},
  two/.style={dash pattern=on 3pt off 2pt,line width=0.9pt}]
  \begin{scope}[xshift=-4.2cm]
    \draw[one,->] (-1.1,0.82)--(-0.15,0.16);
    \draw[one,->] (-1.1,-0.82)--(-0.15,-0.16);
    \draw[two] (-0.15,0)--(0.78,0);
    \draw[one,->] (0.78,0.16)--(1.75,0.82);
    \draw[one,->] (0.78,-0.16)--(1.75,-0.82);
    \node at (0.28,0.36) {\(\phi_2\)};
    \node at (0,-1.18) {\(s\)};
  \end{scope}
  \node at (-1.82,0) {\(+\)};
  \begin{scope}
    \draw[one,->] (-1.1,0.82)--(-0.18,0.28);
    \draw[one,->] (-0.18,0.28)--(1.1,0.82);
    \draw[one,->] (-1.1,-0.82)--(-0.18,-0.28);
    \draw[one,->] (-0.18,-0.28)--(1.1,-0.82);
    \draw[two] (-0.18,0.28)--(-0.18,-0.28);
    \node[right] at (-0.08,0) {\(\phi_2\)};
    \node at (0,-1.18) {\(t\)};
  \end{scope}
  \node at (1.82,0) {\(+\)};
  \begin{scope}[xshift=4.2cm]
    \draw[one,->] (-1.1,0.82)--(-0.12,0.12);
    \draw[one,->] (-1.1,-0.82)--(-0.12,-0.12);
    \draw[two] (-0.12,-0.12)--(0.72,0.58);
    \draw[one,->] (-0.12,0.12)--(1.25,-0.82);
    \draw[one,->] (0.72,0.58)--(1.25,0.82);
    \node at (0.6,0.16) {\(\phi_2\)};
    \node at (0,-1.18) {\(u\)};
  \end{scope}
\end{tikzpicture}
\caption{Tree exchange routings in the cubic scalar model.  Solid directed
lines are external \(\phi_1\) legs and the dashed internal line is the exchanged
\(\phi_2\) field.  The labels \(s,t,u\) specify which Mandelstam invariant flows
through the internal denominator.}
\label{fig:exchange-model-stu-tree-routings}
\end{figure}
With all external momenta on the \(\phi_1\) mass shell, define
\[
  s=-(k_1+k_2)^2,\qquad
  t=-(k_1-k_3)^2,\qquad
  u=-(k_1-k_4)^2 .
\]
Here the metric is mostly plus, \(k_i^2=-m_1^2\), \(k_1,k_2\) are incoming,
and \(k_3,k_4\) are outgoing, as in
Definition~\ref{def:mostly-plus-mandelstam-variables}.  Momentum conservation
gives the Mandelstam constraint
\[
  s+t+u=4m_1^2 .
\]
The tree-level invariant amplitude is
\[
  \mathcal M_{\mathrm{tree}}(s,t,u)
  =
  g^2
  \left(
    \frac{1}{m_2^2-s-\ii0}
    +
    \frac{1}{m_2^2-t-\ii0}
    +
    \frac{1}{m_2^2-u-\ii0}
  \right).
\]

In the center-of-mass frame for equal incoming masses \(m_1\),
\[
  s=E^2,
  \qquad
  t=-\frac{E^2-4m_1^2}{2}(1-z),
  \qquad
  u=-\frac{E^2-4m_1^2}{2}(1+z),
\]
where \(E\) is the total energy and \(z=\cos\theta\). The two-\(\phi_1\)
threshold is
\[
  s=4m_1^2 .
\]

\section{Pole Below a Two-Particle Threshold}

Assume
\[
  0<m_2<2m_1 .
\]
Then the \(s\)-channel pole \(s=m_2^2\) lies below the two-\(\phi_1\)
threshold. It is outside the physical two-\(\phi_1\) scattering region
\(s\ge4m_1^2\), but it is reached by analytic continuation in the invariant
energy variable.

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.95]
  \draw[->,line width=0.75pt] (-0.4,0)--(7.4,0) node[right] {$s$};
  \draw[->,line width=0.75pt] (0,-1.35)--(0,2.05)
    node[above] {$\operatorname{Im}s$};
  \fill[blue!80!black] (2.1,0) circle (2.6pt);
  \node[blue!80!black,below] at (2.1,-0.16) {$m_2^2$};
  \draw[red!75!black,line width=2.0pt] (4.6,0)--(7.05,0);
  \node[red!75!black,below] at (4.6,-0.18) {$4m_1^2$};
  \node[blue!80!black,align=center] at (1.65,1.18)
    {stable\\pole};
  \draw[blue!80!black,->,line width=0.75pt] (1.78,0.82)--(2.04,0.12);
  \node[red!75!black,align=center] at (5.95,0.78)
    {two-particle\\cut};
  \draw[decorate,decoration={brace,amplitude=5pt},gray!70]
    (2.22,0.48)--(4.45,0.48);
  \node[gray!70!black] at (3.35,0.85) {gap};
\end{tikzpicture}
\end{center}

The scattering conclusion is precise: the two-\(\phi_1\) channel couples to a
stable one-particle state of invariant mass \(m_2\). In the present model that
state is represented explicitly by the field \(\phi_2\). In a theory where no
elementary field for that state is introduced, the same below-threshold pole
is the scattering signature of a stable bound state. The pole criterion and
the microscopic origin of the state are separate statements.

\begin{definition}[Stable-channel sheets]
\label{def:stable-channel-sheet-convention}
Fix a connected stable-channel scattering amplitude, with the identity term
and the overall momentum-conservation delta function removed, and let \(s\)
be the invariant energy variable for a two-particle channel whose threshold is
\(s_{\rm th}\).  The \emph{first sheet}, or physical sheet, is the analytic
germ reached from the Feynman boundary prescription whose physical value on
the open two-particle interval is the upper-edge boundary value
\(\mathcal M^{\rm I}(s+\ii0)\), \(s>s_{\rm th}\).  A real point below that
threshold is on the first sheet when it is reached inside this same germ
without crossing the threshold cut.  Near an ordinary square-root threshold,
one may use a local variable such as
\[
  \beta(s)=\sqrt{1-\frac{s_{\rm th}}{s}},
\]
with the branch fixed by \(\beta(s+\ii0)>0\) on the physical interval.
Continuing through the threshold cut to the adjacent germ changes the sign of
this local square root.  That adjacent germ is called the \emph{second sheet}
relative to the specified channel.  With several open channels, a sheet is not
specified by the word ``second'' alone: one must say which channel cuts have
been crossed, or equivalently which local threshold variables have had their
branches changed.

In this terminology, a stable bound state coupled to the channel appears as an
isolated pole on the first sheet below the corresponding threshold.  A
resonance is instead a pole of an analytically continued stable-channel
amplitude on an adjacent nonphysical sheet, not a pole of the physical
real-axis boundary value.  The resonance chapter constructs this second-sheet
continuation explicitly in an exchange model, and the analyticity chapter
later develops the same terminology for general \(2\to2\) amplitudes.
\end{definition}

\begin{remark}[Existence is dynamical input]
The exchange model above contains the state at \(s=m_2^2\) because the
Lagrangian already contains the stable field \(\phi_2\) and the chosen masses
place its pole below the two-\(\phi_1\) threshold.  In a theory without such an
elementary field, the existence of a below-threshold composite pole is a
nonperturbative fact about the spectrum and the connected four-point function,
requiring dynamical spectral input beyond locality, unitarity, and the local
action.  The rest of this chapter uses the conditional statement: if an isolated
stable pole is present in a two-particle channel, then its position, spin, and
factorized residue have the scattering consequences derived below.
For a Bethe--Salpeter kernel, the corresponding converse requires the analytic
Fredholm hypotheses stated later in
Theorem~\ref{thm:bethe-salpeter-fredholm-pole-criterion}; a below-threshold
eigenvalue of an unspecified or truncated kernel is not, by itself, a theorem
that the exact QFT contains a stable particle.
\end{remark}

\begin{remark}[Near-threshold and unitary-limit caveat]
\label{rem:near-threshold-bound-state-poles}
The inequality \(M_B<2m_1\) is the physical location statement: the state lies
below the two-particle continuum.  The pole statement used in this chapter is
stronger.  It assumes an isolated one-particle spectral point of finite
multiplicity, separated from the continuum and from other singularities by a
nonzero gap, with nonzero overlap with the two-particle channel.  Under those
hypotheses the amplitude has a simple first-sheet pole and a factorized
residue.

This hypothesis is not uniform as the binding energy tends to zero.  In a
two-body channel with equal masses, write the relative momentum by
\[
  s=4(m_1^2+k^2),\qquad
  k=\ii\kappa,\qquad
  \kappa>0
\]
for a below-threshold bound state.  The threshold is \(k=0\).  The effective
range expansion of an \(s\)-wave amplitude has the low-energy denominator
\[
  -a^{-1}+\frac12 r_e k^2-\ii k+\cdots ,
\]
so a shallow bound state has \(\kappa\sim a^{-1}\) when the effective range is
finite.  As \(a^{-1}\to0^+\), the pole reaches the branch point at threshold:
there is no fixed neighborhood of the pole disjoint from the two-particle cut,
and the simple-pole expansion in the invariant \(s\)-plane ceases to be the
right local description.  Depending on the side of the continuation one may
obtain a threshold bound state, a virtual state, or a resonance pole on another
sheet.  In three-body unitary physics, Efimov spectra give an even sharper
warning: infinitely many poles may accumulate at threshold, so the isolated
finite-rank pole hypothesis fails outright.

Near-threshold molecular states and unitary-limit systems therefore require a
separate threshold expansion, usually in the momentum variable \(k\) and with
the branch point kept explicit.  The pole factorization below applies to an
isolated stable bound state before this degeneration has occurred; it is not a
claim that every attractive channel near threshold is described by a single
simple pole in \(s\).
\end{remark}

\begin{theorem}[Isolated spectral atom gives a first-sheet pole]
\label{thm:bound-state-spectral-atom-first-sheet-pole}
Let \(A(x)\) and \(B(x)\) be local or smeared composite operators whose
Fourier transforms carry the same conserved quantum numbers.  Assume that, in
the corresponding translation-invariant Hilbert space, the joint
energy-momentum spectrum contains an isolated finite-multiplicity
one-particle mass shell of mass \(M_B\), with spectral projection
\(\Pi_B\), and that this shell is separated by a positive gap from the
continuum in a neighborhood of total momentum \(P\).  Assume also that the
matrix elements of \(A\) and \(B\) on this shell are nonzero and that the
remaining spectral measure is locally finite across \(P^2=-M_B^2\).  Then
the Fourier transform of the connected time-ordered two-point function
\[
  G_{AB}(P)
  =
  \int \dd^4x\,\ee^{\ii P\cdot x}
  \bra{\vac}\operatorname T A(x)B(0)\ket{\vac}_{\rm conn}
\]
has, in the mostly-plus Feynman convention of this volume, the local form
\[
  G_{AB}(P)
  =
  \frac{-\ii\,R_{AB}(P)}
       {P^2+M_B^2-\ii0}
  +
  G_{AB}^{\rm reg}(P),
  \label{eq:bound-state-spectral-atom-pole}
\]
where \(G_{AB}^{\rm reg}\) is analytic in a sufficiently small first-sheet
neighborhood of the pole and
\[
  R_{AB}(P)
  =
  \sum_{\lambda=1}^{d_B}
  \bra{\vac}A(0)\ket{B,P,\lambda}
  \bra{B,P,\lambda}B(0)\ket{\vac}.
  \label{eq:bound-state-spectral-residue}
\]
Here \(\lambda\) labels an orthonormal basis of the finite spin/internal
degeneracy space of the isolated shell.
\end{theorem}

\begin{proof}
Insert the spectral resolution of the translation generators between the two
operators in the Wightman function
\(\bra{\vac}A(x)B(0)\ket{\vac}\).  The isolated one-particle shell contributes
\[
  \sum_{\lambda}
  \int\frac{\dd^3\vec p}{(2\pi)^3\,2E_{\vec p}}\,
  \ee^{-\ii p\cdot x}
  \bra{\vac}A(0)\ket{B,p,\lambda}
  \bra{B,p,\lambda}B(0)\ket{\vac},
  \qquad
  E_{\vec p}=\sqrt{|\vec p|^2+M_B^2}.
  \]
The negative-time ordering gives the corresponding opposite-energy boundary
value.  Combining the two time orderings and Fourier transforming gives the
standard scalar Feynman denominator in the chapter convention,
\(-\ii/(P^2+M_B^2-\ii0)\), multiplied by the finite-rank residue
\eqref{eq:bound-state-spectral-residue}.  The gap hypothesis separates this
mass shell from the rest of the spectral support.  The remaining spectral
integral is therefore analytic in a neighborhood of the isolated pole on the
first sheet.  This proves \eqref{eq:bound-state-spectral-atom-pole}.
\end{proof}

\begin{theorem}[LSZ factorization of an isolated bound-state pole]
\label{thm:bound-state-lsz-pole-factorization}
Consider a connected \(2\to2\) amplitude between stable scalar external
particles, and let \(P=k_1+k_2=k_3+k_4\) be the total momentum in the
\(s\)-channel, \(s=-P^2\).  Assume the external one-particle LSZ hypotheses
of Chapter~\ref{ch:lsz-reduction}, and assume that the connected
four-point function has an isolated spectral atom of mass \(M_B<2m\) in this
channel satisfying the hypotheses of
Theorem~\ref{thm:bound-state-spectral-atom-first-sheet-pole}.  Then, in a
first-sheet neighborhood of \(s=M_B^2\) disjoint from the two-particle cut,
the invariant amplitude has the factorized form
\[
  \mathcal M_{fi}(s,z)
  =
  \frac{
    \sum_{\lambda=1}^{d_B}
      \Gamma^{\rm out}_{\lambda}(k_3,k_4)\,
      \Gamma^{\rm in}_{\lambda}(k_1,k_2)}
       {M_B^2-s-\ii0}
  +
  \mathcal M^{\rm reg}_{fi}(s,z).
  \label{eq:bound-state-lsz-pole-factorization}
\]
The functions \(\Gamma^{\rm in/out}_{\lambda}\) are the LSZ-amputated matrix
elements coupling the two external particles to the one-particle vector
\(\ket{B,P,\lambda}\).  They depend on the external channel, normalization of
one-particle states, and finite choice of interpolating fields, but the pole
position \(M_B^2\), the finite-rank factorization, and the spin content of the
residue are invariant scattering data.
\end{theorem}

\begin{proof}
Write the connected four-point function in the \(s\)-channel as a
time-ordered two-point function of two composite channel operators.  For
test functions \(f_{\rm in/out}\) of compact support in the relative
coordinates, define
\[
\begin{aligned}
  A_{\rm out}(X)
  &=
  \int\dd^4r\,f_{\rm out}(r)\,
  O_3\!\left(X+\frac r2\right)
  O_4\!\left(X-\frac r2\right),\\
  A_{\rm in}(Y)
  &=
  \int\dd^4r\,f_{\rm in}(r)\,
  O_1\!\left(Y+\frac r2\right)
  O_2\!\left(Y-\frac r2\right).
\end{aligned}
\]
LSZ amputation is applied after the external one-particle Fourier variables
are restored and the test functions are removed in the usual distributional
order.  Theorem
\ref{thm:bound-state-spectral-atom-first-sheet-pole} gives the isolated
denominator \((P^2+M_B^2-\ii0)^{-1}\) and a finite-rank residue equal to the
product of matrix elements through the spectral projection \(\Pi_B\).

The LSZ operation multiplies the four external legs by the inverse external
two-point denominators and then puts each external momentum on the stable
mass shell.  By assumption the external LSZ residues are finite and nonzero,
and \(M_B^2\) is separated from the external one-particle poles and from the
two-particle threshold.  Thus LSZ changes only the numerator factors: it
amputates the external legs in the two matrix elements and leaves the
internal denominator \(P^2+M_B^2-\ii0=M_B^2-s-\ii0\) untouched.  The finite
rank of \(\Pi_B\) gives the displayed sum over \(\lambda\), and all remaining
spectral contributions are regular in the chosen first-sheet neighborhood.
\end{proof}

\section{One-Dimensional Pole Criterion as Orientation}
\label{sec:bound-state-one-dimensional-orientation}

The elementary one-dimensional Schr\"odinger model is useful here only as an
orientation for the bound-state pole criterion.  Its Hilbert-space conventions
belong to the general discussion of continuous spectra in
Section~\ref{sec:rigged-hilbert-continuous-spectrum}, and its unitary
Schr\"odinger evolution is part of the quantum-mechanical setup of
Chapter~\ref{ch:hamiltonian-evolution-time-sliced-path-integrals}.  The
detailed one-dimensional analytic continuation, including outgoing resonance
poles and the two-sheeted energy plane, is treated in
Section~\ref{sec:resonance-one-dimensional-outgoing-poles}.

For the present stable-bound-state purpose, take a short-range Hamiltonian
\[
  H=-\frac{1}{2\mu}\frac{\dd^2}{\dd x^2}+V(x),
  \qquad
  E_{\mathrm{nr}}=\frac{k^2}{2\mu},
\]
with \(V\) real.  After fixing the real-\(k\) scattering normalization, an
outgoing solution has the asymptotic form
\[
  \psi_k^{\mathrm{out}}(x)
  \sim
  S(k)^{-1}\ee^{-\ii kx}+\ee^{\ii kx}
  \qquad (x\to+\infty).
\]
Analytic continuation to \(k=\ii\alpha\), \(\alpha>0\), gives
\[
  \psi_{\ii\alpha}^{\mathrm{out}}(x)
  \sim
  S(\ii\alpha)^{-1}\ee^{\alpha x}
  +
  \ee^{-\alpha x}.
\]
An \(L^2\) bound-state wave function cannot contain the growing term
\(\ee^{\alpha x}\).  Thus normalizability requires
\[
  S(\ii\alpha)^{-1}=0,
  \qquad
  E_{\mathrm{nr}}=-\frac{\alpha^2}{2\mu},
\]
or equivalently a pole of \(S(k)\) at \(k=\ii\alpha\) on the positive imaginary
\(k\)-axis.  In the relativistic amplitude this statement becomes: a stable
bound state below a two-particle threshold appears as a first-sheet pole below
the corresponding branch cut, with the sheet convention of
Definition~\ref{def:stable-channel-sheet-convention}.

\section{Partial Waves}

For an equal-mass scalar elastic channel in four dimensions, write
\[
  z=\cos\theta,
  \qquad
  s=E^2 .
\]
The underlying partial waves are generalized two-particle asymptotic states.
In the center-of-mass frame, with the \(\delta\)-normalized sharp-momentum
basis,
\[
  \langle \vec k_1,\vec k_2
  |E,\vec P=\vec0,\ell,m_\ell\rangle_\delta
  =
  \delta^{(3)}(\vec k_1+\vec k_2)\,
  \delta(\omega_1+\omega_2-E)\,
  \mathcal N(E)Y_{\ell m_\ell}(\widehat k_1),
\]
where
\[
  \omega_i=\sqrt{|\vec k_i|^2+m^2},
  \qquad
  |\vec k_1|=|\vec k_2|=k,
  \qquad
  E=2\sqrt{k^2+m^2}.
\]
For identical scalar bosons,
\[
  \mathcal N(E)
  =
  \left(\frac{2E}{k\,\omega_1\omega_2}\right)^{1/2},
  \qquad
  \ell\in2\mathbb Z_{\ge0}.
\]
For a distinguishable equal-mass pair, such as a charged scalar and its
antiparticle, the factor of \(2\) under the square root is absent and both
even and odd \(\ell\) occur. The ordered-particle invariant amplitude below
uses the \(16\pi\) normalization fixed in
Chapter~\ref{ch:cross-sections-unitarity}; for an unordered identical-boson
channel, the same physical information is obtained after Bose symmetrization
and restriction to even partial waves, with the \(32\pi\) normalization in
Equation~\eqref{eq:identical-partial-wave-conversion}.

Rotational invariance diagonalizes the elastic single-channel matrix element,
\begin{equation}
  |E,\ell,m_\ell\rangle^{\mathrm{in}}
  =
  S_\ell(E)|E,\ell,m_\ell\rangle^{\mathrm{out}}
  \quad
  \text{in the elastic sector}.
  \label{eq:exchange-elastic-partial-wave-s-matrix}
\end{equation}
When additional channels with the same conserved quantum numbers are open,
Equation~\eqref{eq:exchange-elastic-partial-wave-s-matrix} is the elastic
component of a larger unitary matrix and gives \(|S_\ell(E)|\le1\).

In the elastic region, define partial-wave amplitudes by
\begin{equation}
  \mathcal M(s,z)
  =
  16\pi\sum_{\ell=0}^{\infty}
  (2\ell+1)a_\ell(s)P_\ell(z),
  \label{eq:exchange-partial-wave-expansion}
\end{equation}
equivalently
\begin{equation}
  a_\ell(s)
  =
  \frac{1}{32\pi}\int_{-1}^{1}\dd z\,
  P_\ell(z)\mathcal M(s,z).
  \label{eq:exchange-partial-wave-projection}
\end{equation}
This is the ordered \(16\pi\) convention fixed in
Chapter~\ref{ch:cross-sections-unitarity}; the coefficient \(a_\ell(s)\) here
is the same partial-wave amplitude used there.  The letter \(a\) is deliberate:
\(M_\delta\) remains reserved for the \(\delta^{(d)}\)-normalized
sharp-momentum kernel of
Chapter~\ref{ch:cross-sections-unitarity}, whereas \(a_\ell\) is a scalar
coordinate of the invariant amplitude in a fixed angular-momentum channel.
The elastic partial-wave \(S\)-matrix eigenvalue is written
\[
  S_\ell(s)=1+2\ii\,\beta(s)a_\ell(s),
  \qquad
  \beta(s)=\sqrt{1-\frac{4m^2}{s}},
\]
with the branch chosen by \(\beta(s)>0\) for \(s>4m^2\). Elastic unitarity is
\[
  |S_\ell(s)|=1
\]
on the physical elastic interval.

\begin{proposition}[Spin from the angular residue]
\label{prop:bound-state-spin-from-angular-residue}
Let the external particles be spinless and work in the center-of-mass frame.
Suppose the isolated bound-state shell in
Theorem~\ref{thm:bound-state-lsz-pole-factorization} transforms as an
irreducible spin-\(J\) representation of spatial rotations and that the
channel coupling is rotationally covariant.  Then the pole residue in the
scalar \(2\to2\) amplitude is proportional to \(P_J(z)\), with
\(z=\widehat k_{\rm out}\cdot\widehat k_{\rm in}\), and only the partial wave
\(\ell=J\) contains the simple pole.  More precisely,
\[
  \sum_{m=-J}^{J}
  \Gamma^{\rm out}_{Jm}(\widehat k_{\rm out})
  \Gamma^{\rm in}_{Jm}(\widehat k_{\rm in})
  =
  g_{\rm out}g_{\rm in}\,
  \frac{2J+1}{4\pi}\,
  P_J(\widehat k_{\rm out}\cdot\widehat k_{\rm in})
  \label{eq:spin-j-residue-addition-theorem}
\]
after the conventional normalization of spherical harmonics is absorbed into
the constants \(g_{\rm in/out}\).
\end{proposition}

\begin{proof}
For spinless external particles, the angular dependence of a coupling to a
spin-\(J\) intermediate state is an intertwiner from functions on the sphere
to the spin-\(J\) representation.  Rotational covariance fixes this
intertwiner, up to an overall constant, to the spherical harmonics
\[
  \Gamma^{\rm in}_{Jm}(\widehat k)=g_{\rm in}Y_{Jm}(\widehat k),
  \qquad
  \Gamma^{\rm out}_{Jm}(\widehat k)=g_{\rm out}
  \overline{Y_{Jm}(\widehat k)}
  \]
with the conjugation convention determined by in/out orientation.  Summing
over the magnetic quantum number and using the spherical-harmonic addition
theorem gives \eqref{eq:spin-j-residue-addition-theorem}.  Orthogonality of
Legendre polynomials in
\eqref{eq:exchange-partial-wave-projection} then projects this residue only
onto the \(\ell=J\) partial wave.
\end{proof}

A stable spin-\(\ell\) state of mass \(M_B<2m\) coupled to the two-particle
channel therefore gives
\[
  a_\ell(s)
  =
  \frac{r_\ell}{M_B^2-s}
  +\text{regular terms},
\]
where the residue \(r_\ell\) is determined by the squared coupling of the
state to the two-particle channel with angular momentum \(\ell\).

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.95]
  \draw[->,line width=0.75pt] (-0.3,0)--(7.2,0) node[right] {$s$};
  \draw[->,line width=0.75pt] (0,-1.2)--(0,2.05)
    node[above] {\(\operatorname{Im}s\)};
  \draw[red!75!black,line width=2.0pt] (4.75,0)--(6.85,0);
  \node[red!75!black,below] at (4.75,-0.16) {$4m^2$};
  \fill[blue!75!black] (2.55,0) circle (2.5pt);
  \node[blue!75!black,below] at (2.55,-0.16) {$M_B^2$};
  \node[blue!75!black,align=center] at (2.45,1.05)
    {pole in\\\(a_\ell\)};
  \node[red!75!black,align=center] at (5.85,0.82)
    {elastic\\cut};
  \draw[blue!75!black,->,line width=0.8pt] (2.45,0.72)--(2.55,0.12);
\end{tikzpicture}
\end{center}

For the exchange model, the \(s\)-channel term is independent of \(z\). Its
below-threshold pole therefore appears in the \(\ell=0\) partial wave. Higher
spin bound states arise when the pole residue has the angular dependence of
\(P_\ell(z)\), or more generally the corresponding spin-\(\ell\) tensor
structure.

The spin carried by a channel pole can be read directly from this angular
dependence. As an algebraic illustration, consider the numerator structure of
scalar QED, developed later in the gauge-theory chapters; here only the
Lorentz-contracted photon-exchange numerator is used, not the full gauge theory.
The charged scalar-antiscalar annihilation channel has an \(s\)-channel
exchange numerator proportional to
\[
  (p_1-p_2)^\mu(p_3-p_4)_\mu .
\]
In the center-of-mass frame this numerator is
\[
  4k^2\cos\theta
  =
  4k^2P_1(\cos\theta),
\]
up to the overall sign fixed by momentum routing and metric conventions. The
massless pole is therefore an \(\ell=1\) pole in this channel. The invariant
statement is the pole position and spin-one residue in the scalar-antiscalar
scattering amplitude; the microscopic description of the corresponding
one-particle state is additional data of the theory.

\section{Operator Interpretation}

Let \(O_B(x)\) be a local or suitably smeared composite operator with nonzero
overlap with a stable state \(B\),
\[
  \bra{\vac}O_B(0)|B(P)\rangle\ne0,
  \qquad
  P^2=-M_B^2 .
\]
Then the spectral representation of a two-point function of \(O_B\) contains
an atom at \(M_B^2\). The same state appears in four-point functions through a
factorizing pole. For external operators \(O_1,O_2,O_3,O_4\), and
\(P\) the momentum flowing through the channel,
\[
  \widetilde G_4(P;\cdots)
  \sim
  \frac{
    \bra{\vac}O_3O_4|B(P)\rangle
    \langle B(P)|O_1O_2\ket{\vac}}
       {P^2+M_B^2-\ii0}
  +\text{regular terms}.
\]
After LSZ reduction on the external stable particles, this factorization
becomes the bound-state pole of the scattering amplitude.

The particle content of a theory is therefore read from spectral and
asymptotic data. A field appearing in a Lagrangian can interpolate a stable
state, and a composite operator can interpolate a stable state. The analytic
pole and its residue state the channel coupling; additional structural data
state how the corresponding state is represented inside the theory.
