\chapter{\texorpdfstring{Large-\(N\) Two-Dimensional QCD and the Light-Front Bound-State Equation}{Large-N Two-Dimensional QCD and the Light-Front Bound-State Equation}}
\label{ch:large-n-two-dimensional-qcd-light-front}
\ConventionsLorentzian

\section{Regulated Theory and Light-Front Coordinates}

Two-dimensional QCD is a useful solvable gauge-theory laboratory precisely
because the gauge field has no local propagating polarization but still
imposes a nontrivial nonabelian Gauss law.  The model considered in this
chapter has gauge group \(SU(N_c)\), \(N_f\) Dirac fermions in the fundamental
representation, and the same trace convention as the four-dimensional QCD
chapter:
\begin{equation}
\begin{gathered}
  \operatorname{tr}_{\square}(t^at^b)=\delta^{ab},
  \\
  (t^a)^i{}_j(t^a)^k{}_\ell
  =
  \delta^i{}_\ell\delta^k{}_j
  -\frac1{N_c}\delta^i{}_j\delta^k{}_\ell .
\end{gathered}
\label{eq:qcd2-su-n-completeness}
\end{equation}
The second identity is the fundamental completeness relation in this
normalization; it is the same color algebra used in
Chapter~\ref{ch:volume-ii-19}.  The coupling \(g_2\) has mass dimension one,
and the local action is written in the connection normalization
\begin{equation}
  S
  =
  \int\dd^2x\,
  \left[
    -\frac1{4g_2^2}\operatorname{tr}_{\square}(F_{\mu\nu}F^{\mu\nu})
    -\bar q(\gamma^\mu D_\mu+m)q
  \right],
  \qquad
  D_\mu=\partial_\mu-\ii A_\mu .
  \label{eq:qcd2-action}
\end{equation}
For several flavors and masses, \(m\) is replaced by a diagonal mass matrix.
The large-\(N_c\) coupling coordinate used below is
\begin{equation}
  \lambda_2:=g_2^2N_c .
  \label{eq:qcd2-thooft-coupling}
\end{equation}
It has dimensions of mass squared.  Because the generator normalization is
fixed by \eqref{eq:qcd2-su-n-completeness}, no hidden factor of \(1/2\) is
allowed in this definition.

Use light-front coordinates
\begin{equation}
  x^\pm=\frac{x^0\pm x^1}{\sqrt2},
  \qquad
  \dd s^2=-2\,\dd x^+\dd x^- ,
  \label{eq:qcd2-light-front-coordinates}
\end{equation}
and take \(x^+\) as time.  The momentum \(p^+\) is conjugate to \(x^-\), and
a massive on-shell particle has
\[
  2p^+p^-=m^2,\qquad p^+>0 .
\]
The light-front Hamiltonian is \(P^-\), while the invariant mass of a state
with no transverse momentum in two dimensions is
\begin{equation}
  M^2=2P^+P^- .
  \label{eq:qcd2-light-front-mass}
\end{equation}
The general light-front regulator datum, including the mass-shell measure,
the role of positive \(p^+\), and the separate zero-mode/constraint sector,
is Definition~\ref{def:light-front-hamiltonian-regulator-datum}; this
chapter is its large-\(N_c\) two-dimensional gauge-theory specialization.

The manipulations below are made at finite regulator before continuum limits
are discussed.  A convenient regulator is to put \(x^-\) on a circle of
length \(2\pi R\), remove or separately constrain the zero mode, impose a
finite harmonic resolution \(K=RP^+\), and only then study \(K\to\infty\).
Equivalently one may use smooth cutoffs in \(p^+\) and a principal-value
prescription for the instantaneous kernel.  The zero mode of the gauge
constraint is not a harmless detail: it fixes the color-neutral sector and
the global electric-flux data.  The bound-state equation derived below is the
continuum form of the regulated color-singlet two-body problem after this
constraint has been imposed.

\section{Light-Front Gauge Constraint}

Choose light-front gauge
\[
  A_-=0 .
\]
Then \(F_{+-}=-\partial_-A_+\).  The \(A_+\) field is nondynamical; it is the
Lagrange-multiplier remnant of Gauss law written in a gauge where the
constraint can be solved explicitly.  The part of the regulated light-front
action depending on \(A_+\) has the form
\begin{equation}
  S[A_+]
  =
  \int\dd x^+\dd x^-\,
  \left[
    \frac1{2g_2^2}(\partial_-A_+^a)(\partial_-A_+^a)
    +A_+^aJ^{+a}
  \right],
  \label{eq:qcd2-a-plus-action}
\end{equation}
where \(J^{+a}\) is the color current of the dynamical light-front fermion
component.  The sign convention in \eqref{eq:qcd2-a-plus-action} is fixed by
\(\dd s^2=-2\dd x^+\dd x^-\) and by the action
\eqref{eq:qcd2-action}.  Variation of \(A_+^a\) gives
\begin{equation}
  -\frac1{g_2^2}\partial_-^2A_+^a+J^{+a}=0 .
  \label{eq:qcd2-gauss-light-front}
\end{equation}
Let
\[
  G:=(-\partial_-^2)^{-1}
\]
on the zero-mode-free subspace selected by the regulator.  On the infinite
line, \(G\) has kernel \(-|x^-|/2\) as an inverse of \(-\partial_-^2\) on
test functions of total integral zero; on the circle the corresponding
kernel is the periodic zero-average Green function.  Solving
\eqref{eq:qcd2-gauss-light-front} and substituting back gives the positive
instantaneous Coulomb Hamiltonian
\begin{equation}
  P^-_{\rm inst}
  =
  \frac{g_2^2}{2}
  \int\dd x^-\dd y^-\,
  J^{+a}(x^-)\,G(x^--y^-)\,J^{+a}(y^-),
  \label{eq:qcd2-instantaneous-hamiltonian}
\end{equation}
with the zero-mode sector understood as above.

\paragraph{Finite-regulator Gauss-law reduction.}
At finite light-front regulator, after the \(A_+\) zero mode has been either
fixed by the color-neutral Gauss-law constraint or kept as a separate compact
degree of freedom, the nonzero-mode integration over \(A_+\) is an ordinary
Gaussian integration.  The equivalence with the instantaneous Hamiltonian
\eqref{eq:qcd2-instantaneous-hamiltonian} is the following square completion.
On the nonzero-mode regulated space, \(-\partial_-^2\) is a positive
invertible symmetric operator.  Integrating \eqref{eq:qcd2-a-plus-action} by
parts in \(x^-\) gives
\[
  S[A_+]
  =
  \int\dd x^+\,
  \left[
    -\frac1{2g_2^2}\langle A_+,\partial_-^2A_+\rangle
    +\langle A_+,J^+\rangle
  \right].
\]
Completing the square with
\[
  A_+=-g_2^2(-\partial_-^2)^{-1}J^+ + A_+'
\]
separates a Gaussian in \(A_+'\) from the current-current term.  In the
Hamiltonian \(P^-\), the latter appears with the positive kernel \(G\) as in
\eqref{eq:qcd2-instantaneous-hamiltonian}.  The completion of the square does
not determine the zero mode; that mode is exactly the global Gauss-law
constraint and must be part of the regulator datum.

\section{Planar Color-Singlet Mesons}

Consider one quark flavor of mass \(m_1\) and one antiquark flavor of mass
\(m_2\).  A color-singlet light-front meson state of total momentum \(P^+>0\)
has the leading large-\(N_c\) form
\begin{equation}
  \ket{P^+;\phi}
  =
  \frac1{\sqrt{N_c}}
  \sum_{i=1}^{N_c}
  \int_0^1\dd x\,
  \phi(x)\,
  b_i^\dagger(xP^+)d^{\dagger i}((1-x)P^+)\ket{\vac}
  \label{eq:qcd2-meson-state}
\end{equation}
after a choice of light-front creation-operator normalization.  The function
\(\phi\) is a wavefunction on \(0<x<1\).  Its endpoint behavior is part of
the domain of the mass operator.  For positive renormalized endpoint masses,
normalizability and finiteness of the mass expectation force sufficient
vanishing at \(x=0,1\); in the chiral limit an endpoint-constant wavefunction
is allowed by the subtracted Coulomb kernel below.

The free light-front Hamiltonian contributes
\[
  P^-_{\rm kin}
  =
  \frac1{2P^+}
  \left[
    \frac{m_1^2}{x}+\frac{m_2^2}{1-x}
  \right]
\]
on the component with momentum fraction \(x\).  The instantaneous interaction
\eqref{eq:qcd2-instantaneous-hamiltonian} transfers longitudinal momentum
between the quark and antiquark.  The planar color contraction follows from
\eqref{eq:qcd2-su-n-completeness}: the first term preserves the singlet with
coefficient \(N_c\), while the second term is suppressed by \(N_c^{-1}\) in
the normalized singlet matrix element.  The real exchange kernel and the
quark and antiquark self-energy terms combine into a subtracted principal
value operator.

\begin{definition}[Subtracted 't Hooft operator]
\label{def:qcd2-subtracted-thooft-operator}
For \(\gamma_2>0\) and endpoint masses \(m_1^2,m_2^2\geq0\), define
\begin{equation}
  (\mathcal M^2\phi)(x)
  =
  \left[
    \frac{m_1^2}{x}
    +
    \frac{m_2^2}{1-x}
  \right]\phi(x)
  +
  \gamma_2\,
  \operatorname{PV}\int_0^1
  \frac{\phi(x)-\phi(y)}{(x-y)^2}\,\dd y .
  \label{eq:qcd2-subtracted-thooft-operator}
\end{equation}
The principal value is best defined by the quadratic form
\eqref{eq:qcd2-thooft-positive-form}.  For smooth test functions,
it is the finite part obtained by applying the same cutoff to real exchange
and self-energy terms before removing the cutoff.  In the large-\(N_c\) QCD
normalization of this chapter,
\begin{equation}
  \gamma_2=\frac{\lambda_2}{\pi}
  \label{eq:qcd2-thooft-gamma-normalization}
\end{equation}
for the standard continuum principal-value convention.  A finite DLCQ
regulator may instead take \(\gamma_2\) as its displayed matrix coefficient
and use \eqref{eq:qcd2-thooft-gamma-normalization} as the continuum matching
condition.
\end{definition}

\begin{theorem}[Planar light-front meson equation]
\label{thm:qcd2-planar-thooft-equation}
Assume the finite light-front regulator described above, take
\(N_c\to\infty\) with \(\lambda_2=g_2^2N_c\) fixed and \(N_f\) fixed, and
restrict to the color-singlet one-quark-one-antiquark sector with endpoint
masses \(m_1,m_2\) fixed in the chosen renormalization scheme.  If the
regulated planar Hamiltonians converge on this sector to the subtracted
operator of Definition~\ref{def:qcd2-subtracted-thooft-operator}, then the
leading large-\(N_c\) meson masses are the spectral values \(M^2\) satisfying
\begin{equation}
  M^2\phi(x)
  =
  \left[
    \frac{m_1^2}{x}
    +
    \frac{m_2^2}{1-x}
  \right]\phi(x)
  +
  \frac{\lambda_2}{\pi}\,
  \operatorname{PV}\int_0^1
  \frac{\phi(x)-\phi(y)}{(x-y)^2}\,\dd y .
  \label{eq:qcd2-thooft-equation}
\end{equation}
\end{theorem}

\begin{proof}
The free term follows from \(M^2=2P^+P^-\): a quark with momentum \(xP^+\)
contributes \(m_1^2/(2xP^+)\) to \(P^-\), and the antiquark contributes
\(m_2^2/(2(1-x)P^+)\).  Multiplication by \(2P^+\) gives the endpoint mass
operator.

For the interaction, write the current in momentum modes.  At finite
regulator, the kernel \(G=(-\partial_-^2)^{-1}\) contributes a factor
\((k^+)^{-2}\) for a nonzero exchanged longitudinal momentum \(k^+\), with
the zero mode excluded by Gauss law.  Acting on the singlet state
\eqref{eq:qcd2-meson-state}, the color algebra
\eqref{eq:qcd2-su-n-completeness} gives a leading contraction
\(\delta^i{}_\ell\delta^k{}_j\) and a subleading trace subtraction.  Hence the
leading interaction is proportional to \(g_2^2N_c=\lambda_2\).

The real exchange term alone has a nonintegrable \(1/(x-y)^2\) singularity.
The self-energy terms generated by the same instantaneous Hamiltonian are the
terms in which the exchanged longitudinal momentum leaves and returns to the
same fermion line.  With a common regulator for the real and self-energy
pieces, their sum is the finite-part difference
\[
  \operatorname{PV}\int_0^1
  \frac{\phi(x)-\phi(y)}{(x-y)^2}\,\dd y .
\]
The coefficient of this subtracted kernel is \(\lambda_2/\pi\) in the
continuum momentum convention used for light-front creation operators.  The
assumed convergence of the regulated planar Hamiltonians then gives
\eqref{eq:qcd2-thooft-equation}.  The statement is deliberately phrased with
the regulator-convergence hypothesis exposed: the algebra above derives the
finite-regulator reduction and the planar kernel, while the continuum
operator domain is an analytic input.
\end{proof}

\begin{remark}[Relation to the unsubtracted form]
A common finite-part coordinate writes the same regulated planar equation
in the form
\[
  M^2\phi(x)
  =
  \left[
    \frac{\widetilde m_1^2}{x}
    +
    \frac{\widetilde m_2^2}{1-x}
  \right]\phi(x)
  -
  \frac{\lambda_2}{\pi}\,
  \operatorname{PV}\int_0^1
  \frac{\phi(y)}{(x-y)^2}\,\dd y .
\]
It is a different finite-part coordinate for the same planar bound-state
equation.  The singular term proportional to \(\phi(x)\) has been assigned
to the endpoint masses \(\widetilde m_i^2\).  The subtracted form
\eqref{eq:qcd2-thooft-equation} displays the real--self-energy pairing and
has a positive quadratic form, so it is the convention used in this
monograph.
\end{remark}

\section{Quadratic Form, Positivity, and the Chiral Limit}

\paragraph{Positive form of the subtracted kernel.}
For a real smooth function \(\phi\) on \([0,1]\),
\begin{equation}
  \int_0^1\dd x\,\phi(x)
  \operatorname{PV}\int_0^1
  \frac{\phi(x)-\phi(y)}{(x-y)^2}\,\dd y
  =
  \frac12
  \int_0^1\dd x\int_0^1\dd y\,
  \frac{(\phi(x)-\phi(y))^2}{(x-y)^2}.
  \label{eq:qcd2-thooft-positive-form}
\end{equation}
This identity is an elementary symmetrization of the principal-value kernel,
but it is the useful way to remember the sign.  Regulate the singularity by
deleting \(|x-y|<\epsilon\).  The regulated bilinear expression is
\[
  I_\epsilon
  =
  \int_0^1\dd x\int_{|x-y|>\epsilon}\dd y\,
  \phi(x)\frac{\phi(x)-\phi(y)}{(x-y)^2}.
\]
Interchange \(x\) and \(y\) in the same integral.  Since
\((x-y)^2=(y-x)^2\), the average of the two equal expressions is
\[
  I_\epsilon
  =
  \frac12
  \int_{|x-y|>\epsilon}\dd x\,\dd y\,
  \frac{(\phi(x)-\phi(y))^2}{(x-y)^2}.
\]
For smooth \(\phi\), the numerator is \(O((x-y)^2)\) near the diagonal, so
the right hand side has a finite \(\epsilon\downarrow0\) limit.  This proves
\eqref{eq:qcd2-thooft-positive-form}.  Adding the endpoint mass terms gives
nonnegative contributions when \(m_i^2\geq0\).
Thus, for \(m_1^2,m_2^2,\gamma_2\geq0\), the quadratic form of
\(\mathcal M^2\) is nonnegative on its form domain.

In the equal-mass chiral limit \(m_1=m_2=0\), the constant wavefunction
\[
  \phi_0(x)=1
\]
lies in the kernel of the subtracted Coulomb operator.  Therefore
\[
  M_0^2=0
\]
in the leading planar equation.  This massless meson is a special feature of
the large-\(N_c\) two-dimensional model and of the order of limits used here.
General statements about four-dimensional QCD and the Schwinger-model mass
generation of Chapter~\ref{ch:schwinger-model} require their own local gauge
dynamics and anomaly structure.

\section{Endpoint Exponents}
\label{sec:qcd2-endpoint-exponents}

The endpoint behavior of the wavefunction is part of the domain of the
light-front mass operator.  It is fixed by the cancellation of the leading
singular terms in the same finite-part convention that defines
\eqref{eq:qcd2-subtracted-thooft-operator}.

\begin{lemma}[Endpoint finite part]
\label{lem:qcd2-endpoint-finite-part}
Let \(0<\beta<1\), and define
\[
  I_\beta(x)
  :=
  \operatorname{PV}\int_0^1
  \frac{x^\beta-y^\beta}{(x-y)^2}\,\dd y .
\]
As \(x\downarrow0\),
\begin{equation}
  I_\beta(x)
  =
  \bigl(\pi\beta\cot(\pi\beta)-1\bigr)x^{\beta-1}
  +O(1).
  \label{eq:qcd2-endpoint-kernel-asymptotic}
\end{equation}
The analogous right-endpoint statement follows by replacing \(x\) with
\(1-x\).
\end{lemma}

\begin{proof}
Set \(y=xt\).  Then
\[
  I_\beta(x)
  =
  x^{\beta-1}
  \operatorname{PV}\int_0^{1/x}
  \frac{1-t^\beta}{(1-t)^2}\,\dd t .
\]
The tail from \(1/x\) to \(\infty\) contributes \(O(x^{1-\beta})\) to the
integral in \(t\), hence \(O(1)\) to \(I_\beta(x)\).  It remains to compute
the finite part
\[
  C(\beta):=
  \operatorname{PV}\int_0^\infty
  \frac{1-t^\beta}{(1-t)^2}\,\dd t .
\]
Integrating by parts with the pole at \(t=1\) removed symmetrically gives
\[
  C(\beta)
  =
  -1+\beta\operatorname{PV}\int_0^\infty
  \frac{t^{\beta-1}}{1-t}\,\dd t .
\]
The principal-value integral is
\[
  \operatorname{PV}\int_0^\infty
  \frac{t^{\beta-1}}{1-t}\,\dd t
  =
  \int_0^1\frac{t^{\beta-1}-t^{-\beta}}{1-t}\,\dd t
  =
  \psi(1-\beta)-\psi(\beta)
  =
  \pi\cot(\pi\beta),
\]
where \(\psi=\Gamma'/\Gamma\), and the last equality is the logarithmic
derivative of Euler's reflection formula
\(\Gamma(\beta)\Gamma(1-\beta)=\pi/\sin(\pi\beta)\).  Thus
\(C(\beta)=\pi\beta\cot(\pi\beta)-1\), proving
\eqref{eq:qcd2-endpoint-kernel-asymptotic}.
\end{proof}

\begin{lemma}[Endpoint exponents]
\label{lem:qcd2-endpoint-exponents}
Suppose an eigenfunction of
\eqref{eq:qcd2-subtracted-thooft-operator} has nonzero leading behavior
\[
  \phi(x)=c_1x^{\beta_1}+o(x^{\beta_1}),
  \qquad
  \phi(x)=c_2(1-x)^{\beta_2}+o((1-x)^{\beta_2}),
\]
with \(c_1,c_2\ne0\) and \(0\leq\beta_i<1\).  For \(m_i^2>0\), the endpoint
exponents lie in \(0<\beta_i<1\) and obey
\begin{equation}
  \frac{m_i^2}{\gamma_2}
  =
  1-\pi\beta_i\cot(\pi\beta_i),
  \qquad i=1,2.
  \label{eq:qcd2-endpoint-exponent-equation}
\end{equation}
For \(m_i^2=0\), the allowed leading exponent is \(\beta_i=0\) in the
subtracted convention.
\end{lemma}

\begin{proof}
Near \(x=0\), the endpoint mass term in
\eqref{eq:qcd2-subtracted-thooft-operator} gives
\[
  m_1^2c_1x^{\beta_1-1}.
\]
Lemma~\ref{lem:qcd2-endpoint-finite-part} gives the singular part of
the subtracted Coulomb term,
\[
  \gamma_2c_1
  \bigl(\pi\beta_1\cot(\pi\beta_1)-1\bigr)x^{\beta_1-1}.
\]
The eigenvalue term \(M^2\phi(x)\) is only \(O(x^{\beta_1})\).  Therefore the
coefficient of \(x^{\beta_1-1}\) must vanish, giving
\eqref{eq:qcd2-endpoint-exponent-equation} for \(i=1\).  The right endpoint
is identical after \(x\mapsto1-x\).

The function \(F(\beta)=1-\pi\beta\cot(\pi\beta)\) is positive and strictly
increasing from \(0\) to \(+\infty\) on \(0<\beta<1\).  Indeed
\[
  F'(\beta)
  =
  \pi^2\beta\csc^2(\pi\beta)-\pi\cot(\pi\beta)
  =
  \frac{\pi}{\sin^2(\pi\beta)}
  \bigl(\pi\beta-\sin(\pi\beta)\cos(\pi\beta)\bigr)>0,
\]
because \(\pi\beta>\sin(\pi\beta)\cos(\pi\beta)\) for \(0<\beta<1\).  Hence
each positive \(m_i^2/\gamma_2\) determines a unique exponent.  When
\(m_i^2=0\), the same equation has the endpoint solution \(\beta_i=0\), which
is the constant endpoint behavior used in the massless chiral solution above.
\end{proof}

\section{Closed Form Domain and Discrete Spectrum}
\label{sec:qcd2-closed-form-domain}

The endpoint analysis determines the local asymptotics of eigenfunctions, but
it is not by itself a definition of the spectral problem.  The continuum
object is the self-adjoint operator associated with the closed positive form
that was already visible in \eqref{eq:qcd2-thooft-positive-form}.  This is the
coordinate in which the statement ``the spectrum of the 't Hooft equation''
has a precise Hilbert-space meaning.

Let
\[
  \mathcal H=L^2((0,1),\dd x)
\]
and, for \(m_1^2,m_2^2,\gamma_2\geq0\), define
\begin{equation}
\begin{split}
  \mathfrak q_{m_1,m_2,\gamma_2}[\phi]
  &:=
  \int_0^1
  \left(
    \frac{m_1^2}{x}
    +
    \frac{m_2^2}{1-x}
  \right)|\phi(x)|^2\,\dd x
  \\
  &\quad
  +
  \frac{\gamma_2}{2}
  \int_0^1\dd x\int_0^1\dd y\,
  \frac{|\phi(x)-\phi(y)|^2}{(x-y)^2}.
\end{split}
\label{eq:qcd2-closed-form}
\end{equation}
The form domain is
\begin{equation}
  \mathcal Q_{m_1,m_2,\gamma_2}
  =
  \left\{
    \phi\in L^2(0,1)\;:\;
    \mathfrak q_{m_1,m_2,\gamma_2}[\phi]<\infty
  \right\}.
  \label{eq:qcd2-form-domain}
\end{equation}
When an endpoint mass is positive, the corresponding weighted \(L^2\) term in
\eqref{eq:qcd2-closed-form} is part of the definition: for example
\(m_1^2>0\) requires
\(\int_0^1x^{-1}|\phi(x)|^2\dd x<\infty\).  When the endpoint mass is zero,
that endpoint weight is absent.  This is why the constant wavefunction is
admissible in the double chiral limit but not for a positive endpoint mass.

\begin{proposition}[Self-adjoint 't Hooft mass operator]
\label{prop:qcd2-self-adjoint-thooft-operator}
For \(m_1^2,m_2^2\geq0\) and \(\gamma_2>0\), the form
\(\mathfrak q_{m_1,m_2,\gamma_2}\) on
\(\mathcal H=L^2(0,1)\) is densely defined, nonnegative, and closed.  Hence
there is a unique nonnegative self-adjoint operator
\(\mathcal M^2_{m_1,m_2,\gamma_2}\) such that
\[
  \mathfrak q_{m_1,m_2,\gamma_2}[\eta,\phi]
  =
  \langle \eta,\mathcal M^2_{m_1,m_2,\gamma_2}\phi\rangle_{L^2}
\]
for all \(\eta\in\mathcal Q_{m_1,m_2,\gamma_2}\) and all
\(\phi\) in the operator domain.  On smooth functions for which the displayed
principal value exists and the endpoint terms are integrable, this operator
acts by
\[
  (\mathcal M^2_{m_1,m_2,\gamma_2}\phi)(x)
  =
  \left(
    \frac{m_1^2}{x}
    +
    \frac{m_2^2}{1-x}
  \right)\phi(x)
  +
  \gamma_2\operatorname{PV}\int_0^1
  \frac{\phi(x)-\phi(y)}{(x-y)^2}\,\dd y .
\]
Its resolvent is compact.  Therefore its spectrum consists of eigenvalues of
finite multiplicity, accumulating only at \(+\infty\).  If
\(m_1=m_2=0\), the constant functions are precisely the zero eigenspace.  If
at least one endpoint mass is positive, zero is not an eigenvalue.
\end{proposition}

\begin{proof}
The domain is dense because \(C_c^\infty(0,1)\) lies in
\(\mathcal Q_{m_1,m_2,\gamma_2}\) and is dense in \(L^2(0,1)\).  Positivity is
immediate from \eqref{eq:qcd2-closed-form}.  For closedness, let
\((\phi_n)\) be Cauchy in the form norm
\[
  \|\phi\|_{\mathcal Q}^2
  =
  \|\phi\|_{L^2}^2+\mathfrak q_{m_1,m_2,\gamma_2}[\phi].
\]
Then \(\phi_n\to\phi\) in \(L^2\) for some \(\phi\).  Passing to a subsequence
gives almost-everywhere convergence.  Fatou's lemma applied to the endpoint
weights and to the positive measure on
\((0,1)^2\setminus\{x=y\}\) gives
\[
  \mathfrak q_{m_1,m_2,\gamma_2}[\phi]
  \leq
  \liminf_{n\to\infty}
  \mathfrak q_{m_1,m_2,\gamma_2}[\phi_n],
\]
so \(\phi\in\mathcal Q_{m_1,m_2,\gamma_2}\).  Applying the same lower
semicontinuity to \(\phi_n-\phi_m\) and then sending \(m,n\to\infty\) shows
that \(\phi_n\to\phi\) in the form norm.  Thus the form is closed.  The
representation theorem for densely defined closed nonnegative forms gives the
unique nonnegative self-adjoint operator associated with the form.

For a smooth \(\phi\) and a smooth test function \(\eta\), regulate the
singularity by deleting \(|x-y|<\epsilon\).  Symmetrizing the bilinear form
gives
\[
  \frac{\gamma_2}{2}\int_{|x-y|>\epsilon}
  \frac{(\overline{\eta(x)}-\overline{\eta(y)})
        (\phi(x)-\phi(y))}{(x-y)^2}\,\dd x\,\dd y
  =
  \gamma_2\int_0^1\overline{\eta(x)}
  \int_{|x-y|>\epsilon}
  \frac{\phi(x)-\phi(y)}{(x-y)^2}\,\dd y\,\dd x .
\]
The finite part as \(\epsilon\downarrow0\) is exactly the principal-value
operator displayed above.  The endpoint multiplication terms are ordinary
weighted \(L^2\) pairings.  This proves the agreement with the differential
form of the equation on the smooth core where all terms are defined.

It remains to identify the spectral type.  The double integral in
\eqref{eq:qcd2-closed-form} is the homogeneous \(H^{1/2}\)-seminorm on the
interval, up to the harmless normalization convention for the finite
interval.  More explicitly, extending \(\phi\) to a function on a slightly
larger interval by a bounded extension operator and then periodizing on a
circle, the kernel form controls
\[
  \sum_{n\in\mathbb Z}|n|\,|\widehat\phi_n|^2
\]
on the extension, while \(\|\phi\|_{L^2}^2\) controls the \(n=0\) coefficient.
The diagonal argument for Fourier series then gives compactness of the unit
ball of \(H^{1/2}\) in \(L^2\): truncate to \(|n|\leq N\), obtain compactness
in the finite-dimensional truncated space, and make the tail uniformly small
by the \(H^{1/2}\) bound.  The endpoint weights only shrink the form domain,
so the embedding
\(\mathcal Q_{m_1,m_2,\gamma_2}\hookrightarrow L^2(0,1)\) is compact.  A
closed semibounded form with compact form-domain embedding has compact
resolvent: \((\mathcal M^2_{m_1,m_2,\gamma_2}+1)^{-1}\) maps \(L^2\)
boundedly into the form domain and then compactly back to \(L^2\).
Therefore the spectrum is purely discrete with finite multiplicities and no
finite accumulation point.

Finally, \(\mathfrak q[\phi]=0\) forces
\(|\phi(x)-\phi(y)|^2=0\) for almost every pair \((x,y)\), hence
\(\phi\) is almost everywhere constant.  If both endpoint masses vanish,
these constants indeed have zero form value.  If at least one endpoint mass
is positive, a nonzero constant has infinite or positive endpoint-weighted
form value, so only the zero vector can have zero form value.
\end{proof}

\begin{remark}[Spectral coordinates]
\label{rem:qcd2-min-max-spectral-coordinates}
The eigenvalues can be characterized by the Rayleigh--Ritz principle applied
to \(\mathfrak q_{m_1,m_2,\gamma_2}\).  This is the continuum statement
behind the finite DLCQ matrices below: finite-dimensional approximations are
useful only insofar as their forms converge to \eqref{eq:qcd2-closed-form} in
a sense strong enough to control these variational values.  Changing the
endpoint mass coordinate or the zero-mode prescription changes the form and
hence changes the spectral problem being approximated.
\end{remark}

\section{Finite DLCQ Matrix and Numerical Benchmark}

The finite matrix used in
Chapter~\ref{ch:constructive-hamiltonian-truncation-dlcq-benchmarks} is the
DLCQ discretization of the subtracted form above.  Fix \(K\ge3\), grid points
\[
  x_n=\frac nK,\qquad n=1,\ldots,K-1,
\]
and a coefficient \(\gamma_2>0\).  Define
\begin{equation}
  (M_K^2)_{nm}
  =
  \delta_{nm}
  \left[
    \frac{m_1^2}{x_n}
    +
    \frac{m_2^2}{1-x_n}
    +
    \gamma_2K
    \sum_{\substack{j=1\\j\ne n}}^{K-1}\frac1{(n-j)^2}
  \right]
  -
  (1-\delta_{nm})\frac{\gamma_2K}{(n-m)^2}.
  \label{eq:qcd2-thooft-dlcq-matrix}
\end{equation}
Then, for \(v\in\mathbb R^{K-1}\),
\begin{equation}
  v^TM_K^2v
  =
  \sum_{n=1}^{K-1}
  \left[
    \frac{m_1^2}{x_n}
    +
    \frac{m_2^2}{1-x_n}
  \right]v_n^2
  +
  \gamma_2K
  \sum_{1\le n<m\le K-1}
  \frac{(v_n-v_m)^2}{(n-m)^2}.
  \label{eq:qcd2-thooft-dlcq-quadratic-form}
\end{equation}
This identity is the finite-dimensional shadow of
\eqref{eq:qcd2-thooft-positive-form}.  It also fixes the sign of the
off-diagonal matrix elements.  In the massless case, the constant vector is
an exact zero mode of the kernel part at every \(K\), matching the continuum
constant solution.

\begin{controlledapproximation}[DLCQ use of the 't Hooft equation]
A DLCQ calculation of two-dimensional large-\(N_c\) QCD must state the
harmonic resolution \(K\), the treatment of zero modes, the endpoint mass
coordinate, the coefficient \(\gamma_2\) and its matching to
\(\lambda_2/\pi\), the normalization of finite eigenvectors, and the
extrapolation model for \(K\to\infty\).  A finite eigenvalue of
\eqref{eq:qcd2-thooft-dlcq-matrix} is a certified eigenvalue of a regulated
matrix; it becomes a continuum meson mass only after these convergence data
are supplied.
\end{controlledapproximation}

The companion script
\[
  \texttt{calculation-checks/thooft\_model\_checks.py}
\]
verifies the finite color normalization, the quadratic-form identity
\eqref{eq:qcd2-thooft-dlcq-quadratic-form}, positivity for positive endpoint
masses, and the exact massless constant zero mode of the subtracted kernel.

\section{\texorpdfstring{Scope of the Two-Dimensional Large-\(N_c\) Construction}{Scope of the Two-Dimensional Large-Nc Construction}}

The model gives an exact leading-planar bound-state equation in a genuine
confining gauge theory.  Its content is the following precise construction.
\begin{enumerate}
  \item The meson equation follows from a gauge constraint, the color-singlet
  Hilbert-space sector, and planar color algebra.
  \item The absence of transverse gauge polarizations in \(1+1\) dimensions
  makes the gauge field a constraint variable while preserving a nontrivial
  nonabelian Coulomb kernel.
  \item Large-\(N_c\) factorization makes one-meson sectors stable at leading
  order.  Meson interactions and widths are organized by subleading powers of
  \(1/N_c\).
  \item The light-front construction is a statement about a declared
  regulator, a zero-mode prescription, and an endpoint-mass coordinate.
  Changing any of these data changes the operator whose spectrum is being
  studied.
\end{enumerate}
Thus two-dimensional large-\(N_c\) QCD is a benchmark for confinement,
bound-state equations, light-front regularization, and the relation between
gauge constraints and color-singlet Hilbert-space sectors.  Its comparison
with four-dimensional QCD is structural rather than literal: the
four-dimensional theory has propagating gluons, running coupling,
spin-dependent interactions, and multi-particle channels that require the
separate developments of the later QCD chapters.
