\chapter{Lattice Yang--Mills as a Nonperturbative Formulation}
\label{ch:lattice-yang-mills-nonperturbative-definition}
\ConventionsEuclidean

This chapter gives a regulator-level construction of pure Yang--Mills theory
which does not begin by choosing a covariant gauge.  The construction is a
finite-dimensional integral over compact groups at finite lattice spacing and
finite volume.  Its continuum limit is a separate problem: one must prove, or
assume with stated evidence, that a sequence of finite-cutoff theories has
gauge-invariant Schwinger functions with a limit satisfying the Euclidean
axioms and yielding a positive-energy Hilbert-space theory after
Osterwalder--Schrader reconstruction.

\section{Finite Lattice Gauge Fields}

Fix a compact Lie group \(G\).  For most formulas below \(G=SU(N)\), and
\(\operatorname{tr}\) denotes the trace in the fundamental representation with
\(\operatorname{tr}(t^a t^b)=\delta^{ab}\), matching the continuum
Yang--Mills convention of
Chapter~\ref{ch:vol2-classical-yang-mills-matter}.  Let
\[
  \Lambda_{a,L}
  =
  (a\mathbb Z/L\mathbb Z)^4
\]
be a four-dimensional periodic hypercubic lattice, where \(a>0\) is the
lattice spacing and \(L/a\) is an integer.  The positively oriented links are
\[
  E^+(\Lambda_{a,L})
  =
  \{(x,\mu)\mid x\in\Lambda_{a,L},\ \mu=1,\ldots,4\}.
\]
The link \((x,\mu)\) joins \(x\) to \(x+a\widehat\mu\).  A lattice gauge field
is a function
\[
  U:E^+(\Lambda_{a,L})\longrightarrow G,
  \qquad
  (x,\mu)\longmapsto U_\mu(x).
\]
The inverse orientation is assigned the inverse group element,
\[
  U_{-\mu}(x+a\widehat\mu)=U_\mu(x)^{-1}.
\]
Thus a lattice gauge field is a point of the compact manifold
\[
  \mathcal A_{\Lambda}=G^{E^+(\Lambda_{a,L})}.
\]

The lattice gauge group is
\[
  \mathcal G_{\Lambda}=G^{\Lambda_{a,L}}.
\]
An element \(\Omega\in\mathcal G_{\Lambda}\) acts by
\begin{equation}
  U_\mu(x)\longmapsto
  U^\Omega_\mu(x)
  =
  \Omega(x)U_\mu(x)\Omega(x+a\widehat\mu)^{-1}.
  \label{eq:lattice-link-gauge-transformation}
\end{equation}
This convention matches the Hermitian continuum convention of
Chapter~\ref{ch:vol2-classical-yang-mills-matter}.  For a smooth field with
\[
  U_\mu(x)=\exp\bigl(-\ii a A_\mu(x)+O(a^2)\bigr),
\]
the endpoint rule \eqref{eq:lattice-link-gauge-transformation} gives
\[
  A_\mu^\Omega
  =
  \Omega A_\mu\Omega^{-1}
  +
  \ii\,\Omega\partial_\mu\Omega^{-1}
  +O(a).
\]

For \(\mu<\nu\), define the oriented plaquette based at \(x\) by
\begin{equation}
  U_{\mu\nu}(x)
  =
  U_\mu(x)
  U_\nu(x+a\widehat\mu)
  U_\mu(x+a\widehat\nu)^{-1}
  U_\nu(x)^{-1}.
  \label{eq:lattice-plaquette-definition}
\end{equation}
Under \eqref{eq:lattice-link-gauge-transformation},
\begin{equation}
  U_{\mu\nu}(x)\longmapsto
  \Omega(x)U_{\mu\nu}(x)\Omega(x)^{-1}.
  \label{eq:lattice-plaquette-covariance}
\end{equation}
Therefore every class function of \(U_{\mu\nu}(x)\), and in particular
\(\operatorname{Re}\operatorname{tr}U_{\mu\nu}(x)\), is gauge invariant.

Figure~\ref{fig:lattice-plaquette-gauge-covariance} records the oriented
plaquette convention and its based-loop gauge covariance
\eqref{eq:lattice-plaquette-covariance}, the input for the Wilson action.

\begin{figure}[H]
  \centering
  \resizebox{0.82\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,>=Stealth,every node/.style={font=\scriptsize}]
    \coordinate (x) at (0,0);
    \coordinate (xm) at (3.2,0);
    \coordinate (xn) at (0,2.2);
    \coordinate (xmn) at (3.2,2.2);
    \fill[black] (x) circle (2pt);
    \fill[black] (xm) circle (2pt);
    \fill[black] (xn) circle (2pt);
    \fill[black] (xmn) circle (2pt);
    \draw[->,very thick,blue!70!black] (x)--(xm)
      node[midway,below] {\(U_\mu(x)\)};
    \draw[->,very thick,blue!70!black] (xm)--(xmn)
      node[midway,right] {\(U_\nu(x+a\widehat\mu)\)};
    \draw[->,very thick,blue!70!black] (xmn)--(xn)
      node[midway,above] {\(U_\mu(x+a\widehat\nu)^{-1}\)};
    \draw[->,very thick,blue!70!black] (xn)--(x)
      node[midway,left] {\(U_\nu(x)^{-1}\)};
    \node[below left] at (x) {\(x\)};
    \node[below right] at (xm) {\(x+a\widehat\mu\)};
    \node[above left] at (xn) {\(x+a\widehat\nu\)};
    \node[above right] at (xmn) {\(x+a\widehat\mu+a\widehat\nu\)};
    \node[align=center] at (1.6,1.1)
      {\(U_{\mu\nu}(x)\)\\conjugates at \(x\)};

    \begin{scope}[shift={(5.2,0.25)}]
      \node[draw,rounded corners=2pt,inner sep=5pt,align=center] (links)
        at (0,1.45) {link variables\\\(U_\mu(x)\in G\)};
      \node[draw,rounded corners=2pt,inner sep=5pt,align=center] (plaq)
        at (3.1,1.45) {plaquette\\\(U_{\mu\nu}(x)\)};
      \node[draw,rounded corners=2pt,inner sep=5pt,align=center] (act)
        at (6.0,1.45) {class functions\\\(\operatorname{Re}\operatorname{tr}U_{\mu\nu}\)};
      \draw[->,thick] (links)--(plaq);
      \draw[->,thick] (plaq)--(act);
      \node[align=center] at (3.05,0.05)
        {gauge covariance is exact\\at finite lattice spacing};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The plaquette is the elementary lattice curvature.  Link variables
  transform at their endpoints; around a closed elementary square the endpoint
  factors cancel except for conjugation at the basepoint.}
  \label{fig:lattice-plaquette-gauge-covariance}
\end{figure}

\section{Wilson Action and the Finite Haar Integral}

For \(G=SU(N)\) the Wilson action is
\begin{equation}
  S_{W,\Lambda}[U]
  =
  \frac{\beta}{N}
  \sum_{x\in\Lambda_{a,L}}
  \sum_{1\leq\mu<\nu\leq 4}
	  \operatorname{Re}\operatorname{tr}
	  \bigl(\mathbf 1-U_{\mu\nu}(x)\bigr),
	  \qquad
	  \beta=\frac{N}{g_0^2}.
	  \label{eq:lattice-wilson-action}
\end{equation}
Here \(g_0\) is the lattice coupling parameter in the trace-form continuum
normalization
\[
  S_{\mathrm{cont}}[A]
  =
  \frac1{4g_0^2}\int\operatorname{tr}(F_{\mu\nu}F_{\mu\nu}) .
\]
With the more common generator normalization
\(\operatorname{tr}(t^a t^b)=\frac12\delta^{ab}\), the same Wilson action is
often written with \(\beta=2N/g_0^2\).  The difference is a coordinate change in
the generator and coupling normalization, not a different lattice integral.
The integrand
\(\operatorname{Re}\operatorname{tr}(\mathbf 1-U_{\mu\nu})\) is nonnegative,
because each eigenvalue of \(U_{\mu\nu}\) lies on the unit circle and hence
\(\operatorname{Re}\operatorname{tr}U_{\mu\nu}\leq N\).  Gauge invariance of
\(S_{W,\Lambda}\) follows from
\eqref{eq:lattice-plaquette-covariance}.

Let \(\dd U\) be normalized Haar measure on \(G\).  At finite \(a\) and finite
\(L\), the lattice partition function is the finite-dimensional compact
integral
\begin{equation}
  Z_{\Lambda}(\beta)
  =
  \int_{\mathcal A_{\Lambda}}
  \prod_{\ell\in E^+(\Lambda_{a,L})}\dd U_\ell\,
  \exp\{-S_{W,\Lambda}[U]\}.
  \label{eq:lattice-partition-function}
\end{equation}
Since \(\mathcal A_{\Lambda}\) is compact and the integrand is continuous,
\(Z_{\Lambda}(\beta)\) is finite and strictly positive.  For a bounded
gauge-invariant function \(O:\mathcal A_{\Lambda}\to\mathbb C\),
\begin{equation}
  \langle O\rangle_{\Lambda,\beta}
  =
  Z_{\Lambda}(\beta)^{-1}
  \int_{\mathcal A_{\Lambda}}
  \prod_{\ell}\dd U_\ell\,
  O[U]\,\exp\{-S_{W,\Lambda}[U]\}
  \label{eq:lattice-expectation}
\end{equation}
is a completely defined number.  No division by an infinite gauge-group volume
and no Faddeev--Popov determinant enter this finite-cutoff definition.

The relation to the continuum Yang--Mills action is obtained by evaluating the
plaquette on a smooth continuum connection.  Let
\(x_c=x+\frac a2\widehat\mu+\frac a2\widehat\nu\) be the plaquette center.
With
\[
  U_\mu(x)
  =
  \exp\bigl(-\ii a A_\mu(x+\tfrac a2\widehat\mu)+O(a^3)\bigr),
\]
the Baker--Campbell--Hausdorff expansion gives
\begin{equation}
  U_{\mu\nu}(x)
  =
  \exp\bigl(-\ii a^2 F_{\mu\nu}(x_c)+O(a^4)\bigr),
  \label{eq:lattice-plaquette-continuum-expansion}
\end{equation}
where \(F_{\mu\nu}\) is the Hermitian-coordinate curvature of
Chapter~\ref{ch:vol2-classical-yang-mills-matter}.  For \(SU(N)\),
\(\operatorname{tr}F_{\mu\nu}=0\), and hence
\[
  \operatorname{Re}\operatorname{tr}(\mathbf 1-U_{\mu\nu})
  =
  \frac{a^4}{2}\operatorname{tr}(F_{\mu\nu}(x_c)F_{\mu\nu}(x_c))
  +O(a^6).
\]
Substitution in \eqref{eq:lattice-wilson-action} gives
\[
  \sum_{\mu<\nu}\frac{a^4}{2}\operatorname{tr}(F_{\mu\nu}F_{\mu\nu})
  =
  \frac{a^4}{4}
  \sum_{\mu,\nu}\operatorname{tr}(F_{\mu\nu}F_{\mu\nu}),
\]
because \(F_{\nu\mu}=-F_{\mu\nu}\).  Therefore
\begin{equation}
  S_{W,\Lambda}[U(A)]
  =
  \frac{1}{4g_0^2}
  \int_{\mathbb T_L^4}
  \operatorname{tr}(F_{\mu\nu}F_{\mu\nu})\,\dd^4x
  +O(a^2),
  \label{eq:lattice-wilson-continuum-limit-formal}
\end{equation}
for smooth fields on the torus of side \(L\).  Equation
\eqref{eq:lattice-wilson-continuum-limit-formal} is an expansion of the
classical lattice action on smooth configurations.  The quantum continuum
limit also requires control of the measure as \(a\downarrow0\), \(L\) is sent
to the desired infrared limit, and \(\beta\to\infty\) along a trajectory that
keeps a physical scale fixed.

\section{Reflection Positivity and the Transfer Matrix}

The Wilson lattice theory is Euclidean from the start.  To connect it to a
Hilbert-space theory one uses the lattice version of reflection positivity,
the discrete precursor of the Osterwalder--Schrader condition developed in
Chapter~\ref{ch:os-reconstruction}.

Choose the reflection \(\vartheta\) which sends the Euclidean time coordinate
\(x_4\) to \(-x_4\) and reverses the orientation of reflected links.  Let
\(\mathcal B_+\) be the algebra generated by gauge-invariant functions of link
variables whose support lies in the positive-time half lattice.  For
\(F\in\mathcal B_+\), set
\[
  \Theta F[U]=\overline{F[\vartheta U]}.
\]

\begin{theorem}[Osterwalder--Seiler reflection positivity for the Wilson action]
\label{thm:lattice-wilson-reflection-positivity}
On a reflection-invariant finite hypercubic lattice, the Wilson measure
\[
  \dd\mu_{\Lambda,\beta}(U)
  =
  Z_{\Lambda}(\beta)^{-1}
  \prod_{\ell}\dd U_\ell\,\ee^{-S_{W,\Lambda}[U]}
\]
satisfies
\begin{equation}
  \int \Theta F[U]\,F[U]\,\dd\mu_{\Lambda,\beta}(U)\geq0,
  \qquad
  F\in\mathcal B_+.
  \label{eq:lattice-reflection-positivity}
\end{equation}
\end{theorem}

\begin{proof}
The proof is finite-dimensional.  Haar measure is invariant under reflection,
inversion, and left and right multiplication.  The Wilson action splits into a
positive-time part, its reflected part, and plaquette factors crossing the
reflection plane.  A crossing plaquette factor is a continuous central
function of a group element \(V\):
\[
  K_\beta(V)
  =
  \exp\left\{\frac{\beta}{N}
  \operatorname{Re}\operatorname{tr}V\right\}.
\]
Peter--Weyl theory expands it as
\begin{equation}
  K_\beta(V)
  =
  \sum_{\lambda\in\widehat G}c_\lambda(\beta)\chi_\lambda(V),
  \qquad
  c_\lambda(\beta)
  =
  \int_G K_\beta(V)\overline{\chi_\lambda(V)}\,\dd V .
  \label{eq:lattice-peter-weyl-crossing-kernel}
\end{equation}
For \(G=SU(N)\) with the fundamental Wilson kernel this positivity is
explicit.  Write
\[
  K_\beta(V)
  =
  \exp\left[
    \frac{\beta}{2N}
    \bigl(\operatorname{tr}V+\operatorname{tr}V^{-1}\bigr)
  \right]
  =
  \sum_{r,s\geq0}
  \frac{(\beta/2N)^{r+s}}{r!\,s!}
  (\operatorname{tr}V)^r(\operatorname{tr}V^{-1})^s .
\]
The product
\((\operatorname{tr}V)^r(\operatorname{tr}V^{-1})^s\) is the character of
\((\mathbb C^N)^{\otimes r}\otimes(\overline{\mathbb C^N})^{\otimes s}\).
Decomposing this finite-dimensional representation into irreducibles gives
\[
  (\operatorname{tr}V)^r(\operatorname{tr}V^{-1})^s
  =
  \sum_{\lambda\in\widehat{SU(N)}}
  m_{\lambda;r,s}\chi_\lambda(V),
  \qquad
  m_{\lambda;r,s}\in\mathbb Z_{\geq0}.
\]
The exponential series is absolutely and uniformly convergent on the compact
group, so its character coefficients are
\[
  c_\lambda(\beta)
  =
  \sum_{r,s\geq0}
  \frac{(\beta/2N)^{r+s}}{r!\,s!}\,m_{\lambda;r,s}
  \geq0 .
\]
The same argument applies to a Wilson kernel built from a unitary
representation \(R\) and its conjugate, with \(R^{\otimes r}\otimes
\overline R^{\otimes s}\) replacing the fundamental tensor powers.
Matrix-coefficient orthogonality rewrites
each crossing contribution as a finite or absolutely convergent sum of terms
\[
  c_\lambda(\beta)\,
  F_{\lambda,i}[U_+]\,\overline{F_{\lambda,i}[\vartheta U_+]},
  \qquad c_\lambda(\beta)\geq0,
\]
after the links on the reflection plane have been integrated.  Multiplication
by the positive-time Boltzmann weight and summation over \(\lambda,i\) turns
\eqref{eq:lattice-reflection-positivity} into a sum of squared absolute values
with nonnegative coefficients.
\end{proof}

Reflection positivity produces a Hilbert space at finite lattice spacing by
quotienting \(\mathcal B_+\) by the null vectors of
\eqref{eq:lattice-reflection-positivity} and completing.  Translation by one
lattice unit in Euclidean time gives a positive transfer matrix
\[
  T_a=\ee^{-aH_a}.
\]
Thus the finite lattice theory has a genuine Hilbert-space interpretation with
a nonnegative lattice Hamiltonian \(H_a\).  The continuum question is whether a
family of such OS-positive finite-cutoff theories has a scaling limit whose
Schwinger functions satisfy the continuum hypotheses of
Chapter~\ref{ch:os-reconstruction}.

\section{Thermal Center Symmetry and the Polyakov Loop}
\label{sec:lattice-center-polyakov-loop}

The Wilson lattice also gives a precise finite-cutoff formulation of the
thermal center-symmetry order parameter.  This subsection is kinematic: it
defines the symmetry, the charged operator, and the order of limits needed for
spontaneous symmetry breaking.  It does not assume a continuum proof of
confinement or of a deconfinement transition.

Let the Euclidean lattice have the form
\[
  \Lambda_{a,N_s,N_t}
  =
  (a\mathbb Z/N_s a\mathbb Z)^3
  \times
  (a\mathbb Z/N_t a\mathbb Z),
  \qquad
  \beta_{\mathrm T}=aN_t .
\]
The last circle is the thermal circle of inverse temperature
\(\beta_{\mathrm T}\).  To avoid confusion, the Wilson coupling parameter in
\eqref{eq:lattice-wilson-action} will be denoted by
\(\beta_{\rm lat}=N/g_0^2\) in this subsection.

For \(G=SU(N)\), choose a center element
\[
  z_m=\ee^{2\pi\ii m/N}\mathbf 1_N,
  \qquad m\in\mathbb Z_N .
\]
Fix one Euclidean-time slice \(\tau_0\).  The thermal center transformation
acts on temporal links crossing that slice by
\begin{equation}
  U_4(\mathbf x,\tau_0)\longmapsto z_m\,U_4(\mathbf x,\tau_0),
  \qquad
  U_i(\mathbf x,\tau)\longmapsto U_i(\mathbf x,\tau),
  \qquad
  U_4(\mathbf x,\tau\neq\tau_0)\longmapsto U_4(\mathbf x,\tau).
  \label{eq:lattice-center-slice-transformation}
\end{equation}
It is the finite thermal-lattice representative of the electric center
one-form symmetry of
Section~\ref{sec:higher-form-symmetry-data-gauge-theory} after the Euclidean
time direction has been compactified.

\begin{lemma}[Exact center symmetry of pure thermal lattice Yang--Mills]
\label{lem:lattice-center-symmetry}
For pure \(SU(N)\) Wilson lattice gauge theory on
\(\Lambda_{a,N_s,N_t}\), the transformation
\eqref{eq:lattice-center-slice-transformation} preserves the Haar measure and
the Wilson action.  If dynamical matter of \(N\)-ality \(k_\alpha\) is added,
the unbroken subgroup is
\[
  \{m\in\mathbb Z_N:\exp(2\pi\ii m k_\alpha/N)=1
  \hbox{ for every dynamical } \alpha\},
\]
the same subgroup as in \eqref{eq:center-one-form-unbroken-subgroup}.
\end{lemma}

\begin{proof}
  Haar measure is invariant under left multiplication by the fixed group
  element \(z_m\).  A spatial plaquette contains no transformed temporal link.
  A temporal plaquette crossing the slice contains one transformed temporal
  link \(U_4(\mathbf x,\tau_0)\) and one inverse transformed temporal link
  \(U_4(\mathbf x+a\widehat i,\tau_0)^{-1}\).  Because \(z_m\) is central,
  the phases cancel:
  \[
    z_m U_4(\mathbf x,\tau_0)\cdots
    \bigl(z_m U_4(\mathbf x+a\widehat i,\tau_0)\bigr)^{-1}
    =
    z_m U_4(\mathbf x,\tau_0)\cdots
    U_4(\mathbf x+a\widehat i,\tau_0)^{-1}z_m^{-1}.
  \]
  The center factors commute through the remaining links and disappear inside
  \(\operatorname{Re}\operatorname{tr}\).  Thus every plaquette term and the
  product Haar measure are invariant.

  A dynamical matter hopping term in representation \(R_\alpha\) and
  \(N\)-ality \(k_\alpha\) crossing the thermal slice acquires the phase
  \(z_m^{k_\alpha}\).  Such a term is invariant for all dynamical fields
  precisely when the displayed phase is one for every \(R_\alpha\).
\end{proof}

The Polyakov line at spatial position \(\mathbf x\) is the temporal holonomy
around the thermal circle,
\begin{equation}
  \mathcal P(\mathbf x)
  =
  U_4(\mathbf x,0)
  U_4(\mathbf x,a)\cdots
  U_4(\mathbf x,(N_t-1)a).
  \label{eq:lattice-polyakov-holonomy}
\end{equation}
For an irreducible representation \(R\), define the normalized Polyakov loop
\begin{equation}
  P_R(\mathbf x)
  =
  d_R^{-1}\chi_R(\mathcal P(\mathbf x)).
  \label{eq:lattice-polyakov-loop-definition}
\end{equation}
Under a periodic lattice gauge transformation,
\(\mathcal P(\mathbf x)\) is conjugated at the basepoint, so \(P_R(\mathbf x)\)
is gauge invariant.  Under the center transformation
\eqref{eq:lattice-center-slice-transformation},
\begin{equation}
  P_R(\mathbf x)
  \longmapsto
  \exp\left(\frac{2\pi\ii m k_R}{N}\right)P_R(\mathbf x),
  \label{eq:lattice-polyakov-center-charge}
\end{equation}
where \(k_R\) is the \(N\)-ality of \(R\).  The fundamental Polyakov loop is
therefore a charged local operator of the three-dimensional thermal theory.
Equivalently, it is a line operator of the four-dimensional Euclidean theory
wrapping the thermal circle.

Figure~\ref{fig:lattice-center-polyakov-loop} shows the center transformation on
a thermal slice and the resulting \(N\)-ality charge of a Polyakov loop in
representation \(R\).

\begin{figure}[H]
  \centering
  \resizebox{0.90\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,>=Stealth,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small] at (1.55,3.28) {thermal lattice};
      \draw[step=0.72cm,gray!45] (0,0) grid (3.1,2.45);
      \draw[very thick,red!70!black] (0,1.44)--(3.1,1.44);
      \node[red!70!black,fill=white,inner sep=1pt] at (1.55,1.22)
        {center slice};
      \draw[very thick,blue!70!black,->]
        (1.44,0.05)--(1.44,0.68)--(1.44,1.34)--(1.44,2.02)--(1.44,2.38);
      \node[blue!70!black,align=center] at (0.48,2.90)
        {Polyakov\\line};
      \draw[->,blue!70!black] (0.88,2.68)--(1.34,2.34);
      \node[below] at (1.55,-0.18) {spatial direction};
      \node[rotate=90] at (-0.38,1.22) {thermal circle};
    \end{scope}

    \begin{scope}[shift={(5.55,0.15)}]
      \node[draw,rounded corners=2pt,align=center,inner sep=5pt] (sym)
        at (0,1.75)
        {multiply \(U_4(\mathbf x,\tau_0)\)\\by \(z_m\in Z_N\)};
      \node[draw,rounded corners=2pt,align=center,inner sep=5pt] (plaq)
        at (3.35,2.45)
        {plaquette terms\\unchanged};
      \node[draw,rounded corners=2pt,align=center,inner sep=5pt] (poly)
        at (3.35,1.05)
        {\(P_R(\mathbf x)\mapsto z_m^{k_R}P_R(\mathbf x)\)};
      \draw[->,thick] (sym.east)--(plaq.west);
      \draw[->,thick] (sym.east)--(poly.west);
      \node[align=center,text width=5.3cm] at (1.65,0.02)
        {the symmetry is exact only when every dynamical\\field is neutral under this center element};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{Thermal center symmetry on the lattice.  Multiplying all temporal
  links crossing one time slice by a center element leaves plaquettes
  invariant, while a Polyakov loop winding the thermal circle acquires the
  center phase of its representation.}
  \label{fig:lattice-center-polyakov-loop}
\end{figure}

At finite spatial volume and with no explicit center-breaking source,
\[
  \langle P_\square(\mathbf x)\rangle_{\Lambda,\beta_{\rm lat}}=0
\]
in pure \(SU(N)\) Yang--Mills.  Indeed, averaging the expectation value over
the exact \(\mathbb Z_N\) transformations gives
\(\langle P_\square\rangle=z_m\langle P_\square\rangle\) for every \(m\), and
the only complex number with this property for a nontrivial center character
is zero.  A nonzero order parameter is therefore an infinite-volume
spontaneous-symmetry-breaking statement, not a finite-volume expectation at
zero source.

\begin{definition}[Polyakov-loop order parameter]
\label{def:polyakov-loop-order-parameter}
Let
\[
  \overline P_\square
  =
  N_s^{-3}\sum_{\mathbf x}P_\square(\mathbf x).
\]
Choose a small source \(h\) selecting one center sector, for example by
adding
\(-hN_s^3\operatorname{Re}\overline P_\square\) to the lattice action.  The
fundamental Polyakov-loop order parameter is
\begin{equation}
  \ell
  =
  \lim_{h\downarrow0}
  \lim_{N_s\to\infty}
  \left\langle\overline P_\square\right\rangle_{h,N_s,N_t,\beta_{\rm lat}},
  \label{eq:polyakov-loop-order-parameter}
\end{equation}
with the continuum thermal limit, if desired, taken afterward along a scaling
trajectory with \(aN_t=\beta_{\mathrm T}\) fixed.  Center symmetry is
unbroken when \(\ell=0\) and broken when \(\ell\neq0\).
\end{definition}

The order of limits in \eqref{eq:polyakov-loop-order-parameter} is part of
the definition.  Reversing the limits gives zero in a finite volume because
the exact finite-volume measure is center invariant.

The same operator has a static-source interpretation.  After a center sector
has been selected, or with the small source used in
\eqref{eq:polyakov-loop-order-parameter}, insertion of \(P_R(\mathbf x)\) at
finite lattice spacing computes the partition function in the presence of a
static external source in representation \(R\), divided by the vacuum
partition function:
\begin{equation}
  \langle P_R(\mathbf x)\rangle_{\rm bare}
  =
  \exp\{-\beta_{\mathrm T}
  (F_R^{\rm bare}-F_0)\}.
  \label{eq:polyakov-loop-static-source-free-energy}
\end{equation}
The adjective ``bare'' matters.  In the continuum limit a static line has a
self-energy divergence, so a renormalized Polyakov loop is obtained by a
multiplicative line renormalization,
\[
  P_R^{\rm ren}
  =
  Z_R(a)^{N_t}P_R^{\rm bare},
\]
or equivalently by subtracting the static-source self-energy from
\(F_R^{\rm bare}\).  This renormalization changes the numerical value of
\(\ell\), but it does not change the center charge
\eqref{eq:lattice-polyakov-center-charge} or whether \(\ell\) vanishes.

Similarly, the Polyakov-loop two-point function gives the free energy of a
static source and antisource pair:
\begin{equation}
  \langle P_R(\mathbf x)P_{\overline R}(\mathbf y)\rangle
  =
  \exp\{-\beta_{\mathrm T}
  (F_{R\overline R}(|\mathbf x-\mathbf y|)-F_0)\},
  \label{eq:polyakov-loop-pair-free-energy}
\end{equation}
after the same line renormalization.  This correlator is center neutral even
when each individual Polyakov loop is charged.  If the infinite-volume
thermal state satisfies clustering, then a center-broken phase has
\begin{equation}
  \lim_{|\mathbf x-\mathbf y|\to\infty}
  \langle P_\square(\mathbf x)P_{\overline{\square}}(\mathbf y)\rangle
  =
  |\ell|^2 ,
  \label{eq:polyakov-clustering-broken-center}
\end{equation}
whereas a center-symmetric phase has \(\ell=0\).

\begin{definition}[Deconfinement in the center-symmetry sense]
\label{def:center-deconfinement}
For pure \(SU(N)\) Yang--Mills, a thermal phase is center confined when the
thermal center symmetry is unbroken, \(\ell=0\), and center deconfined when
the thermal center symmetry is spontaneously broken, \(\ell\neq0\).  A
deconfinement transition is a nonanalytic change, in the infinite-volume
thermal free energy or its derivatives, across which this center order
parameter changes its symmetry-realization pattern.
\end{definition}

This definition is sharp for pure Yang--Mills and for theories whose
dynamical matter preserves a nontrivial center subgroup.  With dynamical
fundamental quarks the fundamental center symmetry is explicitly broken, so
the fundamental Polyakov loop is no longer a strict order parameter.  It
remains a useful static-source observable, and rapid finite-temperature
changes in it may diagnose a crossover, but the exact symmetry statement has
been lost.

The center-symmetry order parameter is logically distinct from the
zero-temperature Wilson-loop area law.  The latter concerns large spacelike
or spacetime Wilson loops and static-source potentials at \(T=0\); the former
concerns the realization of the thermal center symmetry and the free energy
of a single static source at \(T>0\).  In a theory where both are under
control they are related by physical continuity, but neither finite-cutoff
definition is a substitute for proving the other's continuum dynamics.

\section{Wilson Loops and the Strong-Coupling Area Law}

Let \(C\) be a closed oriented lattice path.  If \(C\) traverses the oriented
link \(\ell\) with sign \(\epsilon_\ell=\pm1\), set
\[
  U(C)=\prod_{\ell\in C}U_\ell^{\epsilon_\ell},
\]
with the product ordered along the path.  For an irreducible representation
\(R\) of \(G\) with character \(\chi_R\) and dimension \(d_R\), define the
normalized Wilson loop
\begin{equation}
  W_R(C)
  =
  d_R^{-1}\chi_R(U(C)).
  \label{eq:lattice-wilson-loop-definition}
\end{equation}
Gauge factors at adjacent endpoints cancel around the closed curve, so
\(W_R(C)\) is gauge invariant.

The strong-coupling expansion is the character expansion of each plaquette
Boltzmann factor.  Write
\begin{equation}
  \exp\left\{\frac{\beta}{N}
  \operatorname{Re}\operatorname{tr}U_P\right\}
  =
  c_0(\beta)
  \left[
    1+\sum_{\lambda\neq 0}d_\lambda u_\lambda(\beta)\chi_\lambda(U_P)
  \right],
  \qquad
  u_\lambda(\beta)=O(\beta^{n_\lambda}),
  \label{eq:lattice-character-expansion}
\end{equation}
where \(n_\lambda>0\) for every nontrivial representation.  The Haar
orthogonality relation
\begin{equation}
  \int_G
  R^\lambda(U)_{ij}R^\mu(U^{-1})_{kl}\,\dd U
  =
  \frac{\delta_{\lambda\mu}}{d_\lambda}
  \delta_{il}\delta_{jk}
  \label{eq:lattice-haar-orthogonality}
\end{equation}
implies a local constraint: after integrating a link, the representation
matrices carried by the plaquettes and by the Wilson loop must combine to the
trivial representation on that link.

For a fundamental Wilson loop in \(SU(N)\), the lowest nonzero contribution is
therefore obtained by filling a surface \(\Sigma\) with boundary \(C\) by
fundamental plaquettes.  If \(A_{\min}(C)\) is the minimal number of plaquettes
in such a surface, then
\begin{equation}
  \langle W_{\square}(C)\rangle_{\Lambda,\beta}
  =
  K_C\,
  u_{\square}(\beta)^{A_{\min}(C)}
  \bigl(1+O(\beta)\bigr)
  \label{eq:lattice-strong-coupling-leading-area}
\end{equation}
for sufficiently small \(\beta\), with \(K_C>0\) determined by boundary group
integrals and by the normalization of \(W_{\square}\).  More precisely, the
polymer expansion obtained from \eqref{eq:lattice-character-expansion}
converges absolutely when \(\beta\) is small enough; its polymer activities
obey a bound of the form
\[
  \sum_{\Gamma\ni P}|z(\Gamma)|\ee^{\alpha|\Gamma|}
  \leq \alpha
\]
for a positive \(\alpha\), after decreasing \(\beta\) if necessary.  This
Kotecky--Preiss type estimate gives constants \(C(\beta)\) and
\(\sigma(\beta)>0\) such that
\begin{equation}
  |\langle W_{\square}(C)\rangle_{\Lambda,\beta}|
  \leq
  C(\beta)\,
  \exp\{-\sigma(\beta)A_{\min}(C)\}
  \label{eq:lattice-strong-coupling-area-bound}
\end{equation}
uniformly in the finite volume, away from boundary effects.

For a rectangular loop with spatial length \(R=na\) and Euclidean time length
\(T=ma\), \eqref{eq:lattice-strong-coupling-area-bound} gives
\[
  \langle W_{\square}(C_{R,T})\rangle
  \sim
  \exp\{-\sigma_{\rm lat}(\beta)nm\}
  =
  \exp\{-[\sigma_{\rm lat}(\beta)R/a]\,T/a\}.
  \]
The transfer-matrix interpretation therefore gives a static-source potential
which grows linearly with separation in the strong-coupling lattice theory.
This is a rigorous confinement diagnostic at finite lattice cutoff for small
\(\beta\).  The four-dimensional continuum confinement problem asks for
control along a scaling trajectory with \(\beta\to\infty\).

\section{Scaling Trajectories and Dimensional Transmutation}

The continuum limit is not obtained by fixing \(\beta\) and sending
\(a\downarrow0\).  One must choose a trajectory \(g_0(a)\), equivalently
\(\beta(a)=N/g_0(a)^2\) in the trace normalization of this chapter, so that
dimensionless lattice observables scale in
a way compatible with finite physical observables.  For example, if the
continuum theory has a string tension \(\sigma_{\rm phys}\), a scaling
trajectory would satisfy
\[
  a^2\sigma_{\rm lat}(\beta(a))
  \longrightarrow
  a^2\sigma_{\rm phys}
  \quad\text{with}\quad
  a^2\sigma_{\rm phys}\downarrow0.
\]
Equivalently, the correlation length in lattice units must diverge.

Perturbation theory around the Gaussian ultraviolet fixed point predicts the
large-\(\beta\) relation between \(a\) and the lattice coupling.  For pure
\(SU(N)\), define \(\beta_0,\beta_1\) by
\begin{equation}
  \mu\frac{\dd g}{\dd\mu}
  =
  -\beta_0 g^3-\beta_1 g^5+O(g^7),
  \qquad
  \beta_0=\frac{11N}{48\pi^2},
  \qquad
  \beta_1=\frac{34N^2}{3(16\pi^2)^2}.
  \label{eq:lattice-pure-ym-beta-coefficients}
\end{equation}
Then the associated lattice scale \(\Lambda_L\) is characterized by
\begin{equation}
  a\Lambda_L
  =
  \exp\left\{-\frac{1}{2\beta_0g_0^2}\right\}
  (\beta_0g_0^2)^{-\beta_1/(2\beta_0^2)}
  \bigl(1+O(g_0^2)\bigr),
  \qquad
  g_0^2=\frac{2N}{\beta}.
  \label{eq:lattice-asymptotic-scaling}
\end{equation}
The numerical value of \(\Lambda_L\) depends on the lattice action and on the
definition of the bare coupling.  Ratios between \(\Lambda_L\) and scales in
other schemes, such as \(\Lambda_{\overline{\mathrm{MS}}}\), are determined by
matching a short-distance observable in the overlap of the two regularized
perturbation theories.

The formula \eqref{eq:lattice-asymptotic-scaling} gives the perturbative
tangent to the scaling trajectory near \(g_0=0\).  A construction of
four-dimensional Yang--Mills theory must also show that the gauge-invariant
Schwinger functions or an equivalent algebraic object converge along such a
trajectory and that the reconstructed Hilbert-space theory has the desired
spectral properties.

\section{Local Operators, Counterterms, and Lattice Symmetries}

The Wilson lattice action has exact gauge invariance, lattice translations,
hypercubic rotations and reflections, and charge conjugation when the
\(\theta\)-angle is set to zero.  The continuum local terms that can be
generated by the lattice regulator must respect these symmetries.  This gives
the Symanzik form of a local long-distance effective action:
\begin{equation}
  S_{\rm Sym}
  =
  \int \dd^4x\,
  \left[
    \mathcal E_0(a)
    +
    \frac{1}{4g_R(a)^2}
    \operatorname{tr}(F_{\mu\nu}F_{\mu\nu})
    +
    \sum_i a^{d_i-4}c_i(a)\mathcal O_i(x)
  \right]
  +
  \ii\theta\,\frac{1}{8\pi^2}\int\operatorname{tr}(F\wedge F).
  \label{eq:lattice-symanzik-effective-action}
\end{equation}
Here \(\mathcal E_0(a)\) is a vacuum-energy density, \(d_i>4\), and
\(\mathcal O_i\) are gauge-invariant local operators compatible with the
lattice symmetries.  In pure \(SU(N)\) Yang--Mills, with CP imposed, the only
nontrivial gauge-invariant dimension-four term is
\(\operatorname{tr}(F_{\mu\nu}F_{\mu\nu})\).  If CP is not imposed, the
topological density \(\operatorname{tr}(F\wedge F)\) supplies the separate
\(\theta\)-parameter.  Gauge invariance forbids a gluon mass term, and
hypercubic symmetry makes the coefficient of \(F_{\mu\nu}F_{\mu\nu}\) the same
for every coordinate two-plane on an isotropic lattice.

\begin{definition}[Tuning for a lattice continuum limit]
A lattice parameter is tuned when it must be chosen as a function of \(a\) in
order to keep a specified renormalized observable finite and nonzero as
\(a\downarrow0\).  Parameters multiplying operators with dimension below four
are relevant at the Gaussian fixed point; parameters multiplying
dimension-four operators are marginal at tree level and must be classified by
the beta function and by symmetry.
\end{definition}

For pure CP-even \(SU(N)\) Yang--Mills on an isotropic Wilson lattice, this
power-counting and symmetry analysis leaves one running parameter, \(g_0(a)\).
The mass gap, glueball spectrum, and string tension are then outputs of the
continuum theory, if the continuum limit exists.  Different gauge-invariant
lattice actions with the same symmetries change the coefficients of irrelevant
operators in \eqref{eq:lattice-symanzik-effective-action} and change the
relation between their lattice scales.  Symanzik improvement chooses additional
irrelevant plaquette and loop terms so that the leading lattice artifacts begin
at higher powers of \(a\).

For theories whose ultraviolet flow is not asymptotically free, the preceding
one-parameter mechanism may fail to produce an interacting continuum theory.
Four-dimensional scalar \(\phi^4\) theory and four-dimensional QED are the
basic cautionary examples: perturbation theory points toward a Landau-pole
obstruction, and rigorous triviality results constrain large classes of
lattice scalar limits.  Those results do not settle every possible
UV-complete theory that has the same formal perturbative expansion around the
Gaussian point.  The monograph therefore treats triviality and asymptotic
non-freedom as theory-specific continuum-limit questions, not as consequences
of formal power series alone.

\section{Glueball Spectrum and Flux-Tube Observables}
\label{sec:lattice-ym-glueballs-flux-observables}

The preceding sections define finite-regulator expectation values.  To discuss
the pure Yang--Mills spectrum one must say which gauge-invariant Hilbert-space
object is being extracted from those expectations.  At finite spatial volume
and finite lattice spacing, reflection positivity supplies a transfer matrix
\(\mathbb T=\exp(-aH_{\rm lat})\) on the gauge-invariant Hilbert space.  A
glueball is an eigenstate of \(H_{\rm lat}\) in the vacuum superselection
sector, created with nonzero overlap by a gauge-invariant local operator built
from closed Wilson loops or lattice field-strength composites.  This is a
finite-regulator definition.  A continuum glueball mass is a limit of energy
differences along a scaling trajectory, if the limit exists.

Let \(\mathcal O_i(t)\), \(i=1,\ldots,m\), be gauge-invariant operators
projected to a chosen zero-momentum lattice symmetry channel and with vacuum
expectation subtracted when the channel permits it.  The Euclidean correlation
matrix is
\begin{equation}
  C_{ij}(t)
  =
  \langle \mathcal O_i(t)\mathcal O_j(0)\rangle
  -
  \langle \mathcal O_i\rangle\langle\mathcal O_j\rangle .
  \label{eq:lattice-glueball-correlation-matrix}
\end{equation}
If \(0=E_0<E_1\le E_2\le\cdots\) are transfer-matrix energies in this channel,
then the spectral theorem gives
\begin{equation}
  C_{ij}(t)
  =
  \sum_{n\ge1}
  Z_i^{(n)}\overline{Z_j^{(n)}}\,\ee^{-aE_n t},
  \qquad
  Z_i^{(n)}=\bra{\Omega}\mathcal O_i\ket{n}.
  \label{eq:lattice-glueball-spectral-matrix}
\end{equation}
The integer \(t\) is Euclidean time in lattice units.  The connected
subtraction removes the vacuum term; without it the scalar \(0^{++}\) channel
would be dominated by the identity.

The generalized eigenvalue problem used to extract finite-volume energies
from a correlator matrix is a general finite-regulator method, not a special
property of pure Yang--Mills theory.  Its exact finite-state algebra, basis
covariance, and residual criteria are developed in
Section~\ref{sec:lattice-glueball-correlator-gevp}.  In the present chapter
the Yang--Mills-specific content is the construction of gauge-invariant
Wilson-loop source spaces, the cubic-group channel assignment, reflection
positivity of the chosen lattice regulator, and the continuum/infinite-volume
limit that turns finite-regulator energies into pure-glue spectral data.

The continuum spin labels \(J^{PC}\) are recovered only after the cubic
lattice irreducible representations have been compared across lattice
spacings and volumes.  The following table records the operator meaning of the
lowest pure-glue channels; it is not a phenomenological assignment to
QCD-with-quarks resonances.
\[
\begin{array}{c|c|c}
\text{continuum channel} & \text{typical continuum operator} &
\text{cubic-lattice channel} \\
\hline
0^{++} &
\operatorname{tr}(F_{\mu\nu}F_{\mu\nu}) &
A_1^{++} \\
2^{++} &
\operatorname{tr}\!\left(F_{\mu\alpha}F_{\nu\alpha}
-\frac14\delta_{\mu\nu}F_{\rho\sigma}F_{\rho\sigma}\right) &
E^{++}\oplus T_2^{++} \\
0^{-+} &
\operatorname{tr}(F_{\mu\nu}\widetilde F_{\mu\nu}) &
A_1^{-+} \\
0^{+-} &
\text{multi-loop CP-odd scalar combinations} &
A_1^{+-} \\
1^{+-},\,1^{-+} &
\text{hypercubic vector loop combinations} &
T_1^{+-},\,T_1^{-+}
\end{array}
\]
At nonzero lattice spacing the cubic irreducible representation is the exact
label.  A continuum spin claim is supported when the expected degeneracies
among cubic irreducibles emerge under continuum extrapolation; for example a
continuum \(J=2\) state splits into the \(E\) and \(T_2\) representations of
the cubic group at finite \(a\).

\begin{definition}[Pure-glue spectral continuum datum]
\label{def:pure-glue-spectral-continuum-datum}
A pure-glue spectral continuum datum consists of:
\begin{enumerate}
  \item a choice of gauge group and global form;
  \item a reflection-positive lattice action and scaling trajectory;
  \item a basis of gauge-invariant operators in each cubic \(PC\) channel,
  together with smearing or blocking maps stated as regulator definitions;
  \item finite-volume energies extracted from a declared correlator-matrix
  method such as the GEVP of
  Section~\ref{sec:lattice-glueball-correlator-gevp};
  \item continuum and infinite-volume extrapolations of dimensionless ratios
  such as \(M_{J^{PC}}/\sqrt\sigma\) or \(M_{J^{PC}}/\Lambda_{\overline{\rm MS}}\);
  \item a systematic-error budget for excited-state contamination, volume
  effects, lattice-spacing artifacts, topological freezing, and operator-basis
  conditioning.
\end{enumerate}
Only after these data are stated is a numerical glueball spectrum a statement
about pure Yang--Mills theory rather than about a particular finite lattice
calculation.
\end{definition}

The static flux-tube sector is extracted from rectangular Wilson loops or
from Polyakov-line correlators.  For two static sources in a representation
\(R\), the finite-regulator static energy is
\begin{equation}
  E_R(L)
  =
  -\lim_{T\to\infty}\frac1{aT}
  \log
  \langle W_R(L,T)\rangle_{\rm conn},
  \label{eq:lattice-static-energy-from-wilson-loop}
\end{equation}
when the limit exists.  The string tension in a stable center-charge sector is
\[
  \sigma_R=\lim_{L\to\infty}\frac{E_R(L)}{L}.
\]
Chapter~\ref{ch:volume-ii-19} develops the long-string effective expansion:
the Luscher term, excited open and closed flux-tube levels, \(k\)-string
sectors, baryonic flux networks, and the distinction between winding closed
flux tubes and contractible glueball states.  The role of the present chapter
is the regulator-level extraction: the operators are well-defined finite
Haar-integral observables, and the continuum question is whether their
dimensionless spectral data converge.

\begin{controlledapproximation}[Large-\(N\) glueball width counting]
Normalize single-trace glueball operators formally by
\[
  \mathcal O_N(x)=\frac1{N}\operatorname{tr}\mathcal P(F,D)(x),
\]
so that connected two-point functions are \(O(N^0)\) in the confining
large-\(N\) scaling regime.  If the topological large-\(N\) expansion has
isolated
single-glueball poles with LSZ residues of order one, then connected
\(k\)-glueball amplitudes scale as \(N^{2-k}\).  In particular a three-glueball
coupling is \(O(N^{-1})\), and a glueball decay width into two glueballs is
\(O(N^{-2})\), when the decay is kinematically allowed.
The double-line expansion for connected correlators of \(k\) normalized
single-trace insertions has leading sphere topology with \(k\) marked
boundaries, hence \(N^{2-k}\).  Isolated poles with order-one residues do not
change the \(N\)-power when the external legs are amputated.  A decay width is
quadratic in the on-shell three-point coupling and includes no compensating
positive power of \(N\) in the density of a fixed number of color-singlet
final states, so it is \(O(N^{-2})\).
\end{controlledapproximation}

None of the preceding statements proves the four-dimensional continuum
existence and mass-gap theorem for Yang--Mills theory.  They give the precise
finite-regulator and continuum-extrapolation language in which the expected
mass gap, glueball spectrum, and confining flux-string data are formulated.
The rigorous construction problem remains the one stated at the end of this
chapter: produce a continuum local quantum field theory, prove positivity and
locality after reconstruction, and prove a strictly positive lowest
color-singlet excitation above the vacuum.

\section[Theta-Term and Lattice Topology]{The \(\theta\)-Term and Lattice Topology}

For a smooth continuum \(SU(N)\) connection on a closed four-manifold,
\[
  Q[A]
  =
  \frac{1}{8\pi^2}\int \operatorname{tr}(F\wedge F)
  \in\mathbb Z
\]
is the second Chern number.  The Euclidean action may include
\(\ii\theta Q[A]\), and the integer-valuedness of \(Q\) gives
\(2\pi\)-periodicity in \(\theta\).

At finite lattice spacing, a generic lattice gauge field is only an assignment
of group elements to links.  There is no literal smooth bundle connection from
which \(Q\) can be read without an additional construction.  A plaquette
discretization of the density \(\operatorname{tr}(F\wedge F)\) has the correct
classical limit on smooth fields, but it is not integer-valued on arbitrary
lattice fields.  Consequently \(\exp(\ii\theta Q_{\rm plaq})\) need not be
\(2\pi\)-periodic at finite \(a\).

There are several regulator-level ways to define topological charge more
carefully.  A geometric definition interpolates admissible lattice fields to a
bundle over the torus and evaluates the corresponding Chern number.  With
Ginsparg--Wilson fermions, the index of the lattice Dirac operator gives an
integer
\[
  Q_{\rm GW}
  =
  \operatorname{Tr}\,\gamma_5\left(1-\frac{aD_{\rm GW}}{2}\right)
\]
under the usual admissibility hypotheses.  Gradient-flow definitions smooth
the gauge field before evaluating a lattice density and are designed to have a
controlled continuum limit.  These choices agree in the continuum scaling
regime when the required estimates hold, but they are distinct finite-cutoff
definitions.  Thus the \(\theta\)-angle is an additional parameter whose
nonperturbative lattice definition requires its own regulator convention.

\section{Fermions: Doubling and Chiral Lattice Actions}

A vectorlike gauge theory with fermions requires a lattice Dirac operator
coupled to the link variables.  The naive free lattice Dirac operator is
\begin{equation}
  D_{\rm n}(p)
  =
  \frac{\ii}{a}\sum_{\mu=1}^4\gamma_\mu\sin(ap_\mu),
  \qquad
  p_\mu\in[-\pi/a,\pi/a).
  \label{eq:lattice-naive-dirac-symbol}
\end{equation}
Its zeros occur whenever every \(ap_\mu\) is \(0\) or \(\pi\).  In four
dimensions this gives \(2^4=16\) light species near the corners of the
Brillouin zone.  Figure~\ref{fig:lattice-fermion-doubling} displays the
two-dimensional Brillouin-torus version, where edge identifications turn nine
drawn zeros into four inequivalent continuum cones.

\begin{figure}[H]
  \centering
  \resizebox{0.72\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,>=Stealth,every node/.style={font=\scriptsize}]
    \draw[->] (-3.2,0)--(3.62,0) node[right] {\(ap_1\)};
    \draw[->] (0,-3.2)--(0,3.35) node[above] {\(ap_2\)};
    \draw (-3.1416,-0.08)--(-3.1416,0.08) node[below=4pt] {\(-\pi\)};
    \draw (3.1416,-0.08)--(3.1416,0.08) node[below=4pt] {\(\pi\)};
    \draw (-0.08,-3.1416)--(0.08,-3.1416) node[left=4pt] {\(-\pi\)};
    \draw (-0.08,3.1416)--(0.08,3.1416) node[left=4pt] {\(\pi\)};
    \draw[blue!55!black,thick] (-3.1416,-3.1416) rectangle (3.1416,3.1416);
    \draw[dotted,line width=0.7pt] (-3.1416,3.1416)--(3.1416,3.1416);
    \draw[dotted,line width=0.7pt] (-3.1416,-3.1416)--(3.1416,-3.1416);
    \draw[dotted,line width=0.7pt] (-3.1416,-3.1416)--(-3.1416,3.1416);
    \draw[dotted,line width=0.7pt] (3.1416,-3.1416)--(3.1416,3.1416);
    \fill[red!75!black] (0,0) circle (3.0pt);
    \node[above right] at (0,0) {\((0,0)\)};
    \foreach \x in {-3.1416,3.1416} {
      \draw[red!75!black,line width=0.9pt,fill=white] (\x,0) rectangle ++(-0.08,0.08);
      \draw[red!75!black,line width=0.9pt,fill=white] (\x,0) ++(-0.08,-0.08) rectangle ++(0.16,0.16);
    }
    \node[below right] at (3.1416,-0.18) {\((\pi,0)\)};
    \foreach \y in {-3.1416,3.1416} {
      \draw[red!75!black,line width=0.9pt,fill=white]
        (-0.10,\y-0.08)--(0.10,\y-0.08)--(0,\y+0.10)--cycle;
    }
    \node[above right] at (0.12,3.1416) {\((0,\pi)\)};
    \foreach \x in {-3.1416,3.1416} {
      \foreach \y in {-3.1416,3.1416} {
        \draw[red!75!black,line width=1.0pt] (\x-0.11,\y-0.11)--(\x+0.11,\y+0.11);
        \draw[red!75!black,line width=1.0pt] (\x-0.11,\y+0.11)--(\x+0.11,\y-0.11);
      }
    }
    \node[above right] at (3.1416,3.1416) {\((\pi,\pi)\)};
    \foreach \x/\y in {0.36/-0.34,2.72/-0.52,0.36/2.60,2.72/2.60} {
      \begin{scope}[shift={(\x,\y)},scale=0.24]
        \draw[line width=0.75pt] (-0.7,-0.35)--(0,0.7)--(0.7,-0.35);
        \draw[line width=0.75pt] (-0.45,-0.35)--(0,0.25)--(0.45,-0.35);
      \end{scope}
    }
    \node[align=center] at (0,-3.85)
      {edge identifications of the Brillouin torus group the nine displayed
       zeros into\\
       four inequivalent two-dimensional Dirac species; four dimensions have
       \(2^4\) species};
  \end{tikzpicture}
  }
  \caption{The naive lattice derivative vanishes at every Brillouin-zone
  corner.  In two dimensions the center, the two identified horizontal-edge
  zeros, the two identified vertical-edge zeros, and the four identified
  corners give four inequivalent continuum Dirac cones.}
  \label{fig:lattice-fermion-doubling}
\end{figure}

\begin{theorem}[Degree obstruction for chiral lattice symbols]
\label{thm:lattice-degree-obstruction-chiral-symbol}
Let \(f=(f_1,\ldots,f_D):\mathbb T^D\to\mathbb R^D\) be a smooth periodic map
whose zeros are isolated and nondegenerate, and let
\[
  D_f(p)=\ii\sum_{\mu=1}^D\gamma_\mu f_\mu(p)
\]
with Euclidean gamma matrices satisfying
\(\{\gamma_\mu,\gamma_\nu\}=2\delta_{\mu\nu}\).  This symbol anticommutes with
the fixed chirality matrix \(\gamma_{D+1}\).  Assign to a zero \(p_*\) the
chirality sign
\[
  \chi(p_*)=
  \operatorname{sign}\det
  \left(\frac{\partial f_\mu}{\partial p_\nu}(p_*)\right).
\]
Then
\[
  \sum_{f(p_*)=0}\chi(p_*)=0 .
\]
Consequently a chirally symmetric periodic symbol of this form cannot contain
only one nondegenerate continuum Weyl cone.
\end{theorem}

\begin{proof}
For a regular value \(y\in\mathbb R^D\) of \(f\), define the signed preimage
count
\[
  I(y)
  =
  \sum_{f(p)=y}
  \operatorname{sign}\det
  \left(\frac{\partial f_\mu}{\partial p_\nu}(p)\right).
\]
Sard's theorem makes the regular values dense.  On a connected open set of
regular values, \(I(y)\) is locally constant: if \(y(s)\) is a small path of
regular values and \(f(p_0)=y(0)\), the inverse function theorem gives a
unique smooth branch \(p(s)\) with \(f(p(s))=y(s)\), and the determinant sign
cannot change along the branch because it never vanishes.

The image \(f(\mathbb T^D)\) is compact, so there exists a regular value
\(y_\infty\) outside the image; for it \(I(y_\infty)=0\).  Join \(0\) to
\(y_\infty\) by a smooth path and perturb it, with endpoints fixed, so that it
meets the critical-value set only in isolated birth--death crossings.  At such
a crossing the local normal form creates or annihilates two simple preimages
with opposite determinant signs.  Hence the signed count is unchanged across
the crossing.  Therefore \(I(0)=I(y_\infty)=0\), which is the displayed
chirality-sum identity.

Near a nondegenerate zero, \(D_f(p)\) linearizes to
\(\ii\gamma_\mu A^\mu{}_\nu(p-p_*)^\nu\).  The sign of \(\det A\) is exactly
the orientation, hence chirality, of that continuum cone relative to the
chosen gamma-matrix orientation.  A single Weyl cone would have total chirality
\(\pm1\), contradicting the zero sum.
\end{proof}

The full Nielsen--Ninomiya theorem treats broader local,
translation-invariant, chirally symmetric lattice Dirac operators by replacing
\(f/|f|\) with the appropriate normalized symbol on the Brillouin torus with
small neighborhoods of the zeros removed.  The obstruction is the same degree
identity: a compact momentum torus has no boundary, so the local winding
numbers of all zeros must sum to zero.  Later lattice-fermion chapters spell
out how Wilson and Ginsparg--Wilson constructions evade different hypotheses.

Standard lattice fermion constructions evade different hypotheses of this
theorem.
\begin{description}[leftmargin=2.9em,style=nextline]
\item[Wilson fermions]
Add \(r a\,\Delta/2\), giving doublers masses of order \(1/a\).  The Wilson
term breaks chiral symmetry and produces additive mass renormalization, so the
chiral limit requires tuning \(m_0(a)\) to a critical value \(m_c(a)\).
\item[Staggered fermions]
Diagonalize part of the spin structure and reduce the number of tastes.  The
residual taste symmetry and rooting questions must be treated as finite-cutoff
regularization issues.
\item[Domain-wall fermions]
Place chiral modes on separated boundaries of an auxiliary fifth direction.
Finite fifth-dimensional size produces residual chiral-symmetry breaking.
\item[Overlap and Ginsparg--Wilson fermions]
Replace \(\{\gamma_5,D\}=0\) by
\(\gamma_5D+D\gamma_5=aD\gamma_5D\), giving an exact modified lattice chiral
symmetry and an index theorem at finite cutoff under admissibility hypotheses.
\end{description}
For vectorlike QCD, these formulations provide regulator choices for
gauge-invariant correlation functions.  Their continuum equivalence requires
the same kind of scaling-limit control as in the pure gauge theory, together
with the relevant mass and chiral-symmetry tunings.

\section{Status of the Four-Dimensional Continuum Construction}

\begin{hypothesis}[Four-dimensional lattice Yang--Mills continuum limit]
\label{hyp:lattice-yang-mills-continuum-limit}
Let \(G=SU(N)\) and let the Wilson lattice measures
\(\dd\mu_{\Lambda,\beta}\) be defined by
\eqref{eq:lattice-partition-function}.  There exists a scaling trajectory
\(\beta(a)\to\infty\) and an infinite-volume limit such that the
gauge-invariant Schwinger functions of local and loop observables converge,
after the required field and operator normalizations, to Schwinger functions
satisfying the continuum Osterwalder--Schrader hypotheses.  The reconstructed
Hilbert-space theory has a unique vacuum and a positive mass gap.
\end{hypothesis}

This hypothesis is the constructive content of four-dimensional pure
Yang--Mills theory.  The finite lattice theory is rigorous; the hypothesis is
the assertion that the rigorous finite-cutoff objects have the required
continuum limit and spectral gap.  Renormalization-group analyses of lattice
gauge theory, including Ba{\l}aban's multiscale program, prove substantial
estimates toward such a construction.  Other constructive results in related
models and dimensions give additional control of particular mechanisms.  A
complete proof of Hypothesis~\ref{hyp:lattice-yang-mills-continuum-limit} for
four-dimensional nonabelian Yang--Mills theory remains the Yang--Mills
existence and mass-gap problem.

The strong-coupling area law is a theorem inside the small-\(\beta\) lattice
regime.  The weak-coupling continuum regime lies at \(\beta\to\infty\).  For
four-dimensional pure \(SU(2)\) and \(SU(3)\) Wilson gauge theory, numerical
and analytic evidence support a confining continuum limit connected to the
strong-coupling phase without a physical deconfining transition at zero
temperature.  For other groups and lattice actions, bulk lattice-artifact
transitions may occur and must be distinguished from continuum physics.  The
logical role of the lattice construction is therefore precise: it supplies a
finite, gauge-invariant nonperturbative regulator and a clear continuum-limit
problem; it does not by itself solve the mass-gap problem.

Covariant gauge fixing, Faddeev--Popov ghosts, BRST cohomology, and the BV
formalism developed in the following chapters are indispensable perturbative
and cohomological tools.  Their nonperturbative identification with the
continuum theory constructed from the lattice rests on two further inputs:
first, the lattice continuum limit in
Hypothesis~\ref{hyp:lattice-yang-mills-continuum-limit}; second, agreement of
short-distance gauge-invariant observables with the covariant perturbative
expansion order by order after matching the renormalized coupling and local
operator normalizations.
