\chapter{The 1PI Renormalization Group}
\label{ch:volume-ii-1pi-renormalization-group}
\ConventionsEuclidean


\section{Scale-Dependent Coordinates on the 1PI Effective Action}

The previous chapters constructed local counterterms and organized
subdivergence subtractions.  We now use that construction to compare
renormalized 1PI data at different Euclidean momentum scales.

Work in a Euclidean perturbative quantum field theory on \(\R^D\), with a UV
regulator denoted by \(\Lambda\).  The symbol \(\Lambda\) refers to a cutoff
class in Table~\ref{tab:regulator-integration-status-catalog}; if the
calculation is instead made in dimensional regularization, the corresponding
object is a meromorphic formal graph assignment in \(D=d-\varepsilon\), not a
finite-regulator measure.  Let \(\Gamma_\Lambda[\varphi]\) be the regulated
1PI effective action, expressed as a functional of a classical field
\(\varphi\).  Its momentum-space expansion has the form
\[
  \Gamma_\Lambda[\varphi]
  =
  \sum_{n\ge0}\frac1{n!}
  \int
  \prod_{a=1}^n\frac{\dd^D k_a}{(2\pi)^D}\,
  (2\pi)^D\delta\!\left(\sum_{a=1}^n k_a\right)
  \Gamma_\Lambda^{(n)}(k_1,\ldots,k_n)
  \widetilde\varphi(k_1)\cdots\widetilde\varphi(k_n).
\]
The distribution \(\Gamma_\Lambda^{(n)}\) is the regulated 1PI \(n\)-point
vertex.  Momentum conservation allows one momentum argument to be suppressed
when convenient.

A scale-dependent 1PI coordinate is obtained by specifying:
\begin{itemize}
  \item a finite list of local operators \(\mathcal O_I\), each with a number
  \(n_I\) of fields and a derivative tensor structure;
  \item a Euclidean subtraction configuration \(\mathcal S_{I,\mu}\) whose
  momenta have size \(O(\mu)\);
  \item a field normalization convention at the same scale;
  \item a linear projector \(\Pi_I\) extracting the Taylor coefficient of the
  1PI vertex with the tensor structure of \(\mathcal O_I\).
\end{itemize}
The resulting coordinate is denoted \(g_I(\mu)\).  Its precise value depends
on the chosen subtraction configuration and projector; once these choices are
fixed, \(g_I(\mu)\) is a finite renormalized coordinate on the family of
theories under consideration.

The 1PI renormalization group is the differential comparison of these
coordinates at nearby scales:
\[
  \frac{\dd g_I(\mu)}{\dd\log\mu}
  =
  \beta_I(\{g_J(\mu)\};\mu).
\]
The right-hand side is the beta function in the selected coordinate system.
When all dimensionful quantities have been replaced by dimensionless ratios
and included among the coordinates, the vector field on this enlarged
coordinate space has no separate explicit \(\mu\)-argument.

The finite-cutoff Wilsonian effective action uses another construction.  In
that construction one defines an action at cutoff \(\Lambda\) by integrating
over field modes between two cutoffs and studies the dependence of that action
on \(\Lambda\).  That construction will be developed later.  The present
chapter uses the 1PI effective action and subtraction conditions on its
vertices.

At the level of perturbative graphs, the scale comparison is a precise reuse
of lower-order subgraph data.  A one-loop subgraph that determines the
variation of a 1PI coordinate reappears inside higher-order 1PI graphs.  The
renormalization-group equation records this reuse as a differential equation
for the finite coordinate, after the local counterterms have removed the
cutoff dependence.

Figure~\ref{fig:volume-ii-rg-subgraph-reuse} records the graph-theoretic source
of the perturbative RG equation: lower-order local subgraph data reappear inside
higher-order 1PI graphs and therefore control the scale derivative of a finite
coordinate.

\begin{figure}[H]
  \centering
  \resizebox{0.88\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small] at (0,1.65) {tree datum};
      \coordinate (v) at (0,0);
      \foreach \a in {35,145,215,325} {
        \draw[thick] (v)--++({0.95*cos(\a)},{0.95*sin(\a)});
      }
      \fill (v) circle (2pt);
    \end{scope}

    \draw[-{Latex[length=2mm]}, thick] (1.55,0)--(2.25,0);

    \begin{scope}[shift={(3.85,0)}]
      \node[font=\small] at (0,1.65) {one-loop scale dependence};
      \coordinate (a) at (-0.85,0);
      \coordinate (b) at (0.85,0);
      \draw[thick] (-1.8,0)--(a);
      \draw[thick] (b)--(1.8,0);
      \draw[thick] (a)..controls (-0.35,0.72) and (0.35,0.72)..(b);
      \draw[thick] (a)..controls (-0.35,-0.72) and (0.35,-0.72)..(b);
      \fill (a) circle (2pt);
      \fill (b) circle (2pt);
      \draw[orange!85!black, rounded corners, line width=0.9pt]
        (-1.05,-0.88) rectangle (1.05,0.88);
    \end{scope}

    \draw[-{Latex[length=2mm]}, thick] (6.05,0)--(6.75,0);

    \begin{scope}[shift={(8.85,0)}]
      \node[font=\small] at (0,1.65) {same subgraph in a larger 1PI graph};
      \coordinate (a) at (-1.25,0);
      \coordinate (b) at (-0.15,0);
      \coordinate (c) at (1.05,0);
      \draw[thick] (-2.05,0.72)--(a)--(-2.05,-0.72);
      \draw[thick] (1.85,0.72)--(c)--(1.85,-0.72);
      \draw[thick] (a)..controls (-0.9,0.72) and (-0.45,0.72)..(b);
      \draw[thick] (a)..controls (-0.9,-0.72) and (-0.45,-0.72)..(b);
      \draw[thick] (b)..controls (0.25,0.62) and (0.75,0.62)..(c);
      \draw[thick] (b)..controls (0.25,-0.62) and (0.75,-0.62)..(c);
      \draw[orange!85!black, rounded corners, line width=0.9pt]
        (-1.45,-0.88) rectangle (0.05,0.88);
      \fill (a) circle (2pt);
      \fill (b) circle (2pt);
      \fill (c) circle (2pt);
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The scale derivative of a 1PI coordinate is local in perturbation
  theory because the scale-dependent part of a lower-order 1PI subgraph is
  reused inside larger graphs.  The differential equation is written only
  after the subgraph counterterms of the previous chapters have made the
  coordinate comparison finite.}
  \label{fig:volume-ii-rg-subgraph-reuse}
\end{figure}

Figure~\ref{fig:volume-ii-one-pi-rg-coordinate} then identifies the coordinate
comparison itself: a field normalization and a vertex subtraction condition at
scale \(\mu\) define the finite 1PI coordinates transported by the RG vector
field.

\begin{figure}[H]
  \centering
  \resizebox{0.98\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small] at (0,2.25) {field normalization};
      \draw[thick] (-2.2,0.65)--(-0.55,0.65);
      \fill[gray!20,draw=black,thick] (0,0.65) circle (0.55);
      \draw[thick] (0.55,0.65)--(2.2,0.65);
      \node at (0,0.65) {\(\Sigma\)};
      \node[align=center] at (0,-0.45)
        {\(Z_{\rm MOM}(\mu)=\left(1-\partial_{k^2}\Sigma_R\right)^{-1}\)\\
         at \(k^2=\mu^2\)};
      \node[blue!70!black] at (0,-1.35)
        {\([\phi]_\mu=Z_{\rm MOM}(\mu)^{-1/2}\phi_R\)};
    \end{scope}

    \begin{scope}[shift={(5.2,0)}]
      \node[font=\small] at (0,2.25) {1PI vertex coordinate};
      \fill[gray!18, draw=black, thick] (0,0.65) circle (0.48);
      \foreach \a in {35,145,215,325} {
        \draw[thick] ({0.48*cos(\a)},{0.65+0.48*sin(\a)})
          --++({1.0*cos(\a)},{1.0*sin(\a)});
      }
      \node at (0,0.65) {1PI};
      \node[align=center] at (0,-0.55)
        {\(\Pi_I\Gamma^{(n_I)}\) at\\\(\mathcal S_{I,\mu}\)};
      \node[blue!70!black] at (0,-1.35)
        {\(g_I(\mu)\)};
    \end{scope}

    \begin{scope}[shift={(10.2,0)}]
      \node[font=\small] at (0,2.25) {nearby scales};
      \draw[-{Latex[length=2mm]}] (-2.15,-0.8)--(2.25,-0.8)
        node[below] {\(\log\mu\)};
      \draw[-{Latex[length=2mm]}] (-1.9,-1.0)--(-1.9,1.55)
        node[left] {\(g_I\)};
      \draw[blue!70!black, very thick, domain=-1.55:1.55, samples=80]
        plot (\x,{0.1+0.35*exp(0.45*\x)});
      \draw[dashed] (-0.45,-0.8)--(-0.45,0.39);
      \draw[dashed] (0.35,-0.8)--(0.35,0.51);
      \node[below] at (-0.45,-0.87) {\(\mu\)};
      \node[below] at (0.35,-0.87) {\(\mu+\dd\mu\)};
      \draw[<->, orange!85!black, thick] (-0.45,1.08)--(0.35,1.08);
      \node[orange!85!black] at (0,1.33) {\(\beta_I\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The 1PI renormalization group compares coordinate systems on
  renormalized 1PI vertices.  A field normalization is fixed at the same scale
  as the vertex subtraction, and nearby scales are compared by a differential
  equation.}
  \label{fig:volume-ii-one-pi-rg-coordinate}
\end{figure}

\section{Scalar Quartic Theory in Four Euclidean Dimensions}

Let \(D=4\), and consider a real scalar theory with bare Euclidean Lagrangian
\[
  \mathcal L_E
  =
  \frac12(\partial_\mu\phi)(\partial_\mu\phi)
  +
  \frac12 m_0^2\phi^2
  +
  \frac1{4!}g_0\phi^4 .
\]
Choose a renormalized field \(\phi_R=Z_R^{-1/2}\phi\), a renormalized mass
\(m_R\), and a renormalized quartic coupling \(g_R\).  The same bare
Lagrangian can be written as
\[
\begin{aligned}
  \mathcal L_E
  &=
  \frac12(\partial_\mu\phi_R)(\partial_\mu\phi_R)
  +
  \frac12m_R^2\phi_R^2                                      \\
  &\quad
  +(Z_R-1)\left[
    \frac12(\partial_\mu\phi_R)(\partial_\mu\phi_R)
    +
    \frac12m_R^2\phi_R^2
  \right]
  +
  \frac12\delta m^2\phi_R^2
  +
  \frac1{4!}(g_R+\delta g)\phi_R^4 .
\end{aligned}
\]
Here \(Z_R\), \(\delta m^2\), and \(\delta g\) depend on the regulator and on
the chosen renormalization conditions.

Let \(\Gamma_R^{(4)}(k_1,k_2,k_3)\) be the renormalized 1PI four-point vertex,
with \(k_4=-k_1-k_2-k_3\).  A zero-momentum definition such as
\[
  g_R=-\Gamma_R^{(4)}(0,0,0)
\]
is one possible coordinate.  To compare physics at a Euclidean momentum scale
\(\mu\), it is better to define a coordinate using external momenta of size
\(\mu\) and to normalize the external field at the same scale.

The connected two-point function is written in the form
\[
  G_R(k)
  =
  \frac{1}{k^2+m_R^2-\Sigma_R(k)} ,
\]
which defines the self-energy \(\Sigma_R(k)\) in this convention.  The
scale-dependent normalization is
\[
  Z_{\rm MOM}(\mu)
  =
  \left(
    1-\frac{\partial\Sigma_R}{\partial k^2}
  \right)^{-1}_{k^2=\mu^2},
  \qquad
  [\phi]_\mu=Z_{\rm MOM}(\mu)^{-1/2}\phi_R .
\]
Thus the 1PI effective action, when viewed as a functional of
\([\phi]_\mu\), has a canonically normalized quadratic term at momenta
\(k^2=\mu^2\).

For the four-point coordinate, use the symmetric Euclidean subtraction point
\[
  k_i^2=\mu^2,\qquad
  s=t=u=-\frac43\mu^2,
\]
where
\[
  s=-(k_1+k_2)^2,\qquad
  t=-(k_1+k_3)^2,\qquad
  u=-(k_1+k_4)^2 .
\]
These are Euclidean subtraction variables obtained by analytic continuation
from the mostly-plus Mandelstam convention.  With all four Euclidean momenta
taken incoming and \(k_1+k_2+k_3+k_4=0\), the identity fixing the common
channel value is
\[
\begin{aligned}
  s+t+u
  &=
  -\sum_{j=2}^4 (k_1+k_j)^2  \\
  &=
  -\left(
    3k_1^2+\sum_{j=2}^4 k_j^2
    +2k_1\cdot\sum_{j=2}^4 k_j
  \right) \\
  &=
  -\sum_{i=1}^4 k_i^2 .
\end{aligned}
\]
At \(k_i^2=\mu^2\) this gives \(s+t+u=-4\mu^2\); imposing the symmetric
subtraction condition \(s=t=u\) therefore gives \(s=t=u=-4\mu^2/3\).  After
continuing to Lorentzian on-shell scattering and crossing two momenta back to
the physical outgoing convention, the same algebraic relation becomes
\(s+t+u=\sum_i m_i^2\).
Explicitly, if \(p_i\) are mostly-plus Lorentzian momenta and the Euclidean
momenta are continued as \(k_i=(\ii p_i^0,\vec p_i)\), then
\[
  (k_i+k_j)^2_{\rm E}=(p_i+p_j)^2_{\rm L},
  \qquad
  s_{\rm E}:=-(k_1+k_2)^2_{\rm E}
  =
  s_{\rm L}:=-(p_1+p_2)^2_{\rm L}.
\]
The sign flip is between \(s_{\rm E}\) and the bare mostly-plus Lorentzian
quadratic form \((p_1+p_2)^2_{\rm L}\), not between the Euclidean subtraction
variable and the mostly-plus Mandelstam variable used in the scattering
chapters.
The running quartic coupling in this momentum-subtraction scheme is
\[
  g(\mu)
  =
  -Z_{\rm MOM}(\mu)^2
  \Gamma_R^{(4)}(k_1,k_2,k_3)
  \bigg|_{k_i^2=\mu^2,\;s=t=u=-\frac43\mu^2}.
\]

\begin{definition}[Momentum-subtraction quartic coordinate]
For the four-dimensional Euclidean scalar theory above, the momentum-
subtraction quartic coordinate \(g(\mu)\) is the negative of the four-point
1PI vertex at the symmetric Euclidean point, expressed in the field
normalization \([\phi]_\mu\).  The subtraction scheme consists of the
renormalized field convention, the symmetric point, and the sign convention in
the displayed formula.
\end{definition}
In the 1PI vertex convention used here, the tree-level quartic vertex is
\(\Gamma_{\rm tree}^{(4)}=-g_R\).  The explicit minus sign in the definition of
\(g(\mu)\) therefore makes the coordinate agree at tree level with the positive
coefficient multiplying \(\phi^4/4!\) in the Euclidean action.  In
Section~\ref{sec:volume-ii-scalar-quartic-pole-one-loop} the
minimal-subtraction coordinate \(\lambda(\mu)\) is written directly as the
positive coefficient in
\[
  g^\epsilon
  =
  \mu^\epsilon
  \left[
    \lambda(\mu)
    +\frac{3}{16\pi^2}\frac{\lambda(\mu)^2}{\epsilon}
    +\cdots
  \right].
\]
Thus the momentum-subtraction and minimal-subtraction formulas use the same
positive quartic-coordinate orientation; the displayed minus sign appears only
when the coordinate is extracted from the 1PI vertex.

\section{One-Loop Running in the Quartic Example}

At one loop, the four-point bubbles change the quartic vertex, while the
two-point tadpole is independent of \(k^2\).  In the self-energy convention
\[
  G_R(k)=\frac{1}{k^2+m_R^2-\Sigma_R(k)},
\]
the amputated one-loop tadpole is
\[
  \Sigma^{(1)}_{\rm tad}(k)
  =
  -\,\frac{g_R}{2}
  \int^\Lambda\frac{\dd^4\ell}{(2\pi)^4}\,
  \frac{1}{\ell^2+m_R^2},
\]
before the mass counterterm is chosen.  The factor \(1/2\) is the tadpole
symmetry factor.  The loop momentum \(\ell\) closes at the quartic vertex, so no
external momentum flows through the loop and
\[
  \frac{\partial\Sigma^{(1)}_{\rm tad}}{\partial k^2}=0 .
\]
The one-loop subtraction needed in the two-point function is therefore a mass
subtraction, not a kinetic subtraction.  Hence
\[
  Z_R=1+O(g_R^2),
  \qquad
  Z_{\rm MOM}(\mu)=1+O(g_R^2)
\]
for the one-loop computation of \(g(\mu)\).

Figure~\ref{fig:volume-ii-phi-four-one-loop-rg} displays the actual one-loop
quartic coordinate calculation: tree vertex, three bubble channels, and the
local counterterm fixed by the chosen subtraction condition.

\begin{figure}[H]
  \centering
  \resizebox{0.98\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}[shift={(-5.2,0)}]
      \node[font=\small] at (0,1.75) {tree};
      \foreach \a in {45,135,225,315} {
        \draw[thick] (0,0)--++({0.82*cos(\a)},{0.82*sin(\a)});
      }
      \fill[black] (0,0) circle (2pt);
      \node[below] at (0,-1.0) {\(-g_R\)};
    \end{scope}

    \node[font=\large] at (-3.85,0) {\(+\)};

    \begin{scope}[shift={(-2.35,0)}]
      \node[font=\small] at (0,1.75) {\(s\)-channel};
      \coordinate (a) at (-0.55,0);
      \coordinate (b) at (0.55,0);
      \draw[thick] (-1.45,0.68)--(a)--(-1.45,-0.68);
      \draw[thick] (1.45,0.68)--(b)--(1.45,-0.68);
      \draw[thick] (a)..controls (-0.18,0.58) and (0.18,0.58)..(b);
      \draw[thick] (a)..controls (-0.18,-0.58) and (0.18,-0.58)..(b);
      \fill (a) circle (1.8pt);
      \fill (b) circle (1.8pt);
    \end{scope}

    \node[font=\large] at (-0.7,0) {\(+\)};

    \begin{scope}[shift={(0.75,0)}]
      \node[font=\small] at (0,1.75) {\(t\)-channel};
      \coordinate (a) at (0,0.54);
      \coordinate (b) at (0,-0.54);
      \draw[thick] (-1.12,1.1)--(a)--(1.12,1.1);
      \draw[thick] (-1.12,-1.1)--(b)--(1.12,-1.1);
      \draw[thick] (a)..controls (0.58,0.22) and (0.58,-0.22)..(b);
      \draw[thick] (a)..controls (-0.58,0.22) and (-0.58,-0.22)..(b);
      \fill (a) circle (1.8pt);
      \fill (b) circle (1.8pt);
    \end{scope}

    \node[font=\large] at (2.25,0) {\(+\)};

    \begin{scope}[shift={(3.6,0)}]
      \node[font=\small] at (0,1.75) {\(u\)-channel};
      \coordinate (a) at (-0.42,0.42);
      \coordinate (b) at (0.42,-0.42);
      \draw[thick] (-1.25,1.12)--(a)--(-1.25,-1.12);
      \draw[thick] (1.25,1.12)--(b)--(1.25,-1.12);
      \draw[thick] (a)..controls (0.1,0.88) and (0.85,0.1)..(b);
      \draw[thick] (a)..controls (-0.1,-0.88) and (-0.85,-0.1)..(b);
      \fill (a) circle (1.8pt);
      \fill (b) circle (1.8pt);
    \end{scope}

    \node[font=\large] at (5.05,0) {\(+\)};

    \begin{scope}[shift={(6.35,0)}]
      \node[font=\small] at (0,1.75) {counterterm};
      \foreach \a in {45,135,225,315} {
        \draw[thick] (0,0)--++({0.82*cos(\a)},{0.82*sin(\a)});
      }
      \draw[red!75!black, thick] (-0.23,-0.23)--(0.23,0.23);
      \draw[red!75!black, thick] (-0.23,0.23)--(0.23,-0.23);
      \node[red!75!black, below] at (0,-1.0) {\(-\delta g\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The one-loop scalar quartic coordinate receives the tree vertex,
  the three distinct bubble channels, and the local quartic counterterm fixed
  by the chosen subtraction condition.}
  \label{fig:volume-ii-phi-four-one-loop-rg}
\end{figure}

With a momentum cutoff \(\Lambda\), the one-loop relation at the symmetric
subtraction point is
\[
\begin{aligned}
  -g(\mu)
  &=
  -g_R-\delta g
  +
  3\frac{g_R^2}{32\pi^2}
  \int_0^1 \dd x\,
  \left[
    \log\frac{\Lambda^2}{m_R^2-sx(1-x)}-1
  \right]_{s=-\frac43\mu^2}
  +
  O(g_R^3).
\end{aligned}
\]
Choose the finite convention
\[
  g_R=g(0).
\]
This condition fixes \(\delta g\).  Subtracting the \(\mu=0\) equation from
the equation at general \(\mu\) removes the cutoff and gives
\[
  g(\mu)
  =
  g_R
  -
  \frac{3g_R^2}{32\pi^2}
  \int_0^1 \dd x\,
  \log
  \frac{m_R^2}
       {m_R^2+\frac43\mu^2x(1-x)}
  +
  O(g_R^3).
\]
This is a finite relation between two renormalized quartic coordinates in the
same theory.  For \(\mu\gg m_R\), the displayed correction contains a large
logarithm.  A comparison built from nearby scale coordinates is then better
adapted to perturbation theory.

\section{Nearby Scales and the Beta Function}

Let \(\mu'\) be close to \(\mu\).  The same one-loop computation may be
expressed as a relation between the two scale coordinates:
\[
  g(\mu')
  =
  g(\mu)
  -
  \frac{3}{32\pi^2}g(\mu)^2
  \int_0^1 \dd x\,
  \log
  \frac{m_R^2+\frac43\mu^2x(1-x)}
       {m_R^2+\frac43\mu'^2x(1-x)}
  +
  O(g(\mu)^3).
\]
The expansion parameter is now \(g(\mu)\), and the logarithm is small for
\(\mu'\) close to \(\mu\).  Setting \(\mu'=\mu+\dd\mu\) gives
\[
  \frac{\dd g(\mu)}{\dd\log\mu}
  =
  \frac{3}{16\pi^2}g(\mu)^2
  \int_0^1\dd x\,
  \frac{\frac43\mu^2x(1-x)}
       {m_R^2+\frac43\mu^2x(1-x)}
  +
  O(g(\mu)^3).
\]
The integral is a threshold factor.  At scales much larger than the mass,
\(\mu\gg m_R\), it tends to \(1\), and
\[
  \frac{\dd g}{\dd\log\mu}
  =
  \frac{3}{16\pi^2}g^2+O(g^3).
\]
In this scheme and convention, the one-loop beta function is therefore
\[
  \beta(g)=\frac{3}{16\pi^2}g^2+O(g^3)
  \qquad
  (\mu\gg m_R).
\]

The one-loop differential equation can be integrated between two scales:
\[
  \frac1{g(\mu)}-\frac1{g(\mu')}
  =
  -\frac{3}{16\pi^2}\log\frac{\mu}{\mu'} .
\]
Equivalently,
\[
  g(\mu)
  =
  \frac{g(\mu')}
       {1-\frac{3}{16\pi^2}g(\mu')\log(\mu/\mu')} .
\]
This expression contains powers of
\(g(\mu')\log(\mu/\mu')\) to all orders.  Its perturbative domain is controlled
by the running coordinate itself: \(g(\widetilde\mu)\) must remain small for
the scales \(\widetilde\mu\) used in the comparison.

\section{The Landau Scale as a Perturbative Coordinate}

The integrated one-loop equation may be written in terms of a scale
\(\mu_0\) defined by
\[
  \mu_0
  =
  \mu'\exp\!\left[
    \frac{16\pi^2}{3g(\mu')}
  \right].
\]
Then
\[
  g(\mu)
  =
  \frac{1}{\frac{3}{16\pi^2}\log(\mu_0/\mu)} .
\]
For \(0<\mu\ll\mu_0\), the denominator is large and the one-loop coordinate is
small.  As \(\mu\) approaches \(\mu_0\), the perturbative coordinate becomes
large and the one-loop approximation ceases to provide a controlled
description.  Within the one-loop equation, the pair
\((\mu',g(\mu'))\) is equivalently represented by the single scale \(\mu_0\).

Figure~\ref{fig:volume-ii-phi-four-landau-scale} plots this one-loop coordinate
flow and marks the scale \(\mu_0\) at which the perturbative coordinate leaves
the controlled small-coupling region.

\begin{figure}[H]
  \centering
  \resizebox{0.70\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \draw[-{Latex[length=2mm]}] (-3.0,0)--(3.25,0) node[below] {\(\log\mu\)};
    \draw[-{Latex[length=2mm]}] (-2.65,-0.1)--(-2.65,3.35)
      node[left] {\(g(\mu)\)};
    \draw[blue!70!black, very thick, domain=-2.35:2.35, samples=160]
      plot (\x,{0.55/(2.65-\x)});
    \draw[red!75!black, dashed, thick] (2.65,0)--(2.65,3.05);
    \node[red!75!black, above] at (2.65,3.05) {\(\mu_0\)};
    \node[green!45!black, align=center] at (-1.35,0.72)
      {controlled\\\(g\ll1\)};
    \node[orange!85!black, align=center] at (1.15,1.85)
      {large logarithms\\resummed by RG};
    \node[red!75!black, align=left, anchor=west] at (2.82,1.25)
      {large\\coordinate};
  \end{tikzpicture}
  }
  \caption{The one-loop scalar quartic flow grows with the Euclidean momentum
  scale.  The scale \(\mu_0\) is where the one-loop coordinate diverges; the
  controlled perturbative domain is the region where the running coordinate
  remains small.}
  \label{fig:volume-ii-phi-four-landau-scale}
\end{figure}

\begin{remark}[Status of the Landau scale]
The scale \(\mu_0\) is a statement about the one-loop perturbative coordinate.
It gives a useful parametrization of the weak-coupling region and indicates
where that coordinate loses perturbative control.  A statement about the
existence or nonexistence of a continuum scalar theory at arbitrarily short
distances requires additional nonperturbative input.
\end{remark}

\begin{theorem}[Triviality of the standard scalar scaling limits]
\label{thm:standard-four-dimensional-scalar-triviality}
For nearest-neighbor ferromagnetic four-dimensional Ising-type spin systems at
or near the critical point, and for four-dimensional \(\lambda\phi^4\) fields
with a lattice ultraviolet cutoff in the infinite-volume and vanishing-lattice
spacing limit, the scaling limits covered by the
Aizenman--Duminil-Copin theorem are Gaussian.  Equivalently, after the
continuum field normalization, the limiting Schwinger functions satisfy
Wick's rule and the connected \(n\)-point functions with \(n\ge4\) vanish.
Earlier results of Aizenman and Fr\"ohlich establish complementary
triviality, mean-field, and critical-approach statements in \(d>4\) and in
related \(d\ge4\) scalar and Ising-type settings.
\footnote{See M.~Aizenman, ``Proof of the triviality of
\(\phi_d^4\) field theory and some mean-field features of Ising models for
\(d>4\),'' \emph{Phys. Rev. Lett.} \textbf{47} (1981) 1--4,
doi:10.1103/PhysRevLett.47.1; J.~Fr\"ohlich, ``On the triviality of
\(\lambda\phi_d^4\) theories and the approach to the critical point in
\(d\ge4\) dimensions,'' \emph{Nucl. Phys. B} \textbf{200} (1982) 281--296,
doi:10.1016/0550-3213(82)90088-8; M.~Aizenman and H.~Duminil-Copin,
``Marginal triviality of the scaling limits of critical 4D Ising and
\(\phi^4_4\) models,'' \emph{Ann. of Math.} \textbf{194} (2021) 163--235,
doi:10.4007/annals.2021.194.1.3.}
\end{theorem}

The theorem is not a derivation from the one-loop beta function.  Its proofs
use constructive probabilistic representations and correlation inequalities;
in the four-dimensional marginal case the decisive estimate is a logarithmic
improvement of the tree-diagram bound in the random-current representation.
Its role here is to mark the boundary between two questions.  For the
specified positive lattice scalar and Ising-type scaling-limit problems,
removing the cutoff at criticality does not produce a non-Gaussian scalar QFT
in four dimensions.  By contrast, the broader question whether there exists
some ultraviolet-complete
four-dimensional local QFT whose small-coupling expansion agrees with formal
\(\phi^4_4\) perturbation theory to all orders, while differing by effects
invisible to perturbation theory, is not settled by this theorem.

\section{General 1PI Coordinates}

\begin{definition}[Field-normalization factors]
\label{def:three-field-normalization-factors}
Three field-normalization factors occur in this chapter, and they have
different domains of definition.
\begin{center}
\begin{tabular}{p{0.18\textwidth}p{0.72\textwidth}}
\(\text{symbol}\) & meaning \\
\hline
\(Z_R\) &
Cutoff-dependent factor used in the initial renormalization convention, for
example \(\phi_R=Z_R^{-1/2}\phi\) in the scalar quartic example. \\
\(Z_{\rm MOM}(\mu)\) &
Finite momentum-subtraction rescaling between an already renormalized reference
field and the scale-\(\mu\) 1PI coordinate field:
\([\phi]_\mu=Z_{\rm MOM}(\mu)^{-1/2}\phi_R^{\rm ref}\). \\
\(Z_\phi\) &
Total bare-to-renormalized field factor used in the Callan--Symanzik equation
for the selected chart:
\(\phi_R^{\rm chart}=Z_\phi^{-1/2}\phi_0\).
\end{tabular}
\end{center}
Thus \(Z_R\) and \(Z_{\rm MOM}(\mu)\) are not equal as written.  If
\(\phi_R^{\rm ref}=(Z_\phi^{\rm ref})^{-1/2}\phi_0\), then the chart whose
elementary field is \([\phi]_\mu\) has the total factor
\[
  Z_\phi^{\rm chart}
  =
  Z_\phi^{\rm ref}\,Z_{\rm MOM}(\mu).
\]
If the reference field is itself defined by the same momentum-subtraction
condition at \(\mu=\mu_{\rm ref}\), then \(Z_{\rm MOM}(\mu_{\rm ref})=1\); the
cutoff dependence is in \(Z_\phi^{\rm ref}\), not in the finite rescaling.
\end{definition}

Let a regulated Euclidean action be written as
\[
  S_\Lambda[\phi]
  =
  S_{\rm kin,\Lambda}[\phi]
  +
  \int \dd^Dx\,\sum_I g_I^{\rm bare}(\Lambda)\,\mathcal O_I(x).
\]
For each label \(I\), let \(\mathcal O_I\) be a local operator with mass
dimension \(d_I\), field number \(n_I\), and derivative tensor structure
specified by a polynomial \(P_I\) in external momenta.  The corresponding 1PI
coordinate at scale \(\mu\) is extracted from the \(n_I\)-point vertex by
\[
  g_I(\mu)
  =
  \Pi_I
  \left[
    Z_{\rm MOM}(\mu)^{n_I/2}\,
    \Gamma^{(n_I)}(k_1,\ldots,k_{n_I})
  \right]_{\mathcal S_{I,\mu}} .
\]
The projector \(\Pi_I\) includes the sign and tensor normalization
conventions.  In the scalar quartic example, \(\Pi_I\) is multiplication by
\(-1\) at the symmetric four-point configuration.

Two elementary coordinates fix the normalization conventions.  First consider
the two-derivative quadratic operator represented, up to total derivatives and
normalization convention, by \(\mathcal O_{2,2}=\phi\,\Box\phi\), in the same
self-energy convention used above.  Its 1PI Taylor coefficient at
\(k^2=\mu^2\) is
\[
  \Gamma^{(2;2)}\big|_{k^2=\mu^2}
  =
  \left.
  \frac{\partial\Sigma}{\partial k^2}
  \right|_{k^2=\mu^2}
  =
  1-Z_{\rm MOM}(\mu)^{-1}.
\]
With the minus sign included in the quadratic-coordinate projector, the
corresponding renormalized coordinate is
\[
  g_{2,2}(\mu)
  =
  -Z_{\rm MOM}(\mu)\left(1-Z_{\rm MOM}(\mu)^{-1}\right)
  =
  1-Z_{\rm MOM}(\mu).
\]
This is the same field-normalization information expressed as a 1PI
coordinate.  Second, for the mass operator with bare coefficient
\(g_2^{\rm bare}=m_0^2\), the dimensionless coordinate may be written
\[
  \lambda_2(\mu)
  =
  \mu^{-2}Z_{\rm MOM}(\mu)
  \left[
    g_2^{\rm bare}
    -
    \Sigma(k)\big|_{k^2=\mu^2}
  \right],
\]
again in the same self-energy and projector convention.  These examples
make explicit that the field normalization, relevant mass coordinate, and
higher 1PI vertex coordinates are part of one coordinate chart.

\begin{definition}[Mass data and mass coordinates]
\label{def:mass-data-and-mass-coordinates}
Assume that the connected Euclidean two-point function under discussion
satisfies the spectral representation of Chapter~\ref{ch:kallen-lehmann}.  If
the spectral measure has an isolated atom
\[
  \dd\rho_\Phi(s)
  =
  Z_\Phi\,\delta_{m_*^2}+\dd\rho_{\Phi,\rm c}(s),
  \qquad
  Z_\Phi>0,
\]
where \(\delta_{m_*^2}\) is the Dirac measure at \(s=m_*^2\).  If there is no
support below \(m_*^2\) in that channel, then \(m_*\) is the
\emph{spectral pole mass} seen by the interpolating operator \(\Phi\).  If the
corresponding state is a stable one-particle state and another local operator
has nonzero overlap with the same state, it gives the same \(m_*\); the residue
changes with the operator, not the mass.  In a gauge theory this statement is
made for gauge-invariant operators or for physical BRST cohomology states.
Pole positions of gauge-fixed elementary fields are not, by themselves,
physical mass data.

By contrast, \(m_R\) is a finite 1PI subtraction coordinate, and
\(\lambda_2(\mu)\) is a dimensionless running coordinate for the relevant
quadratic deformation.  They are scheme-dependent coordinates on a family of
regulated Green functions.  The lattice parameter \(m_0^2\), or its spin-model
analogue such as \(T-T_c\), is a regulator coordinate that must be tuned as the
cutoff is removed.  None of these three coordinates is a pole mass by
definition.
\end{definition}

\begin{proposition}[Relation of the mass notions]
\label{prop:relation-of-mass-notions}
Suppose the exact renormalized two-point function in the finite 1PI chart has
the form
\[
  G_R(k)
  =
  \frac{1}{k^2+m_R^2-\Sigma_R(k)}
\]
in a neighborhood admitting analytic continuation to the isolated stable pole.
Then the spectral pole mass satisfies
\begin{equation}
  m_*^2
  =
  m_R^2-\Sigma_R(k^2=-m_*^2),
  \label{eq:pole-mass-finite-subtraction-relation}
\end{equation}
with the same sign convention as the displayed Euclidean inverse propagator.
An on-shell mass prescription imposes \(m_R=m_*\) order by order, while a
generic momentum-subtraction or minimal-subtraction prescription does not.  In
an MS chart, \(\lambda_2(\mu)\) is related to \(m_*\) only after a finite
matching condition and RG evolution have been specified.  In a lattice
regularization with lattice spacing \(a\), the bare coordinate \(m_0^2(a)\) is
tuned to a critical surface \(m_{0,c}^2(a)\) for a massless scaling limit, or
away from it so that
\begin{equation}
  m_{\rm gap}
  =
  \lim_{a\to0}\frac{1}{a\,\xi_{\rm lat}(a)}
  \label{eq:lattice-gap-scaling-relation}
\end{equation}
exists.  When the lower edge of the reconstructed spectral support is an
isolated one-particle atom, \(m_{\rm gap}=m_*\); otherwise \(m_{\rm gap}\) is
the threshold edge rather than a pole mass.
\end{proposition}

\begin{proof}
The first assertion is the definition of an isolated K\"allen--Lehmann atom:
near \(s=m_*^2\),
\[
  G_E(k)=\frac{Z_\Phi}{k^2+m_*^2}+\text{term analytic at }k^2=-m_*^2 .
\]
Thus the inverse two-point function vanishes at \(k^2=-m_*^2\).  Substituting
that value into the chosen 1PI inverse-propagator convention gives
\eqref{eq:pole-mass-finite-subtraction-relation}.  A finite change of
renormalization prescription changes the coordinate \(m_R\) and the finite
part of \(\Sigma_R\), but not the spectral singularity of the completed Green
function.

Minimal subtraction fixes a coordinate by pole subtraction in dimensional
regularization.  It contains no condition at the physical pole, so determining
\(m_*\) requires first expressing the finite Green function in that coordinate
and then solving the pole equation.  For the lattice statement, exponential
decay of Euclidean two-point functions gives \(\xi_{\rm lat}^{-1}\) as the
lowest transfer-matrix energy in lattice units in the selected channel.
Multiplication by \(a^{-1}\) converts it to continuum units.
Osterwalder--Schrader reconstruction identifies the limiting transfer-matrix
spectrum with the reconstructed Hamiltonian spectrum under the usual
reflection-positivity and scaling-limit hypotheses.  An isolated lowest
eigenvalue gives the spectral atom and hence \(m_*\); continuous spectral
support beginning at the same energy gives only a gap edge.
\end{proof}

The finite comparison of nearby scales gives
\[
  \frac{\dd g_I(\mu)}{\dd\log\mu}
  =
  \beta_I(\{g_J(\mu)\};\mu).
\]
Since \(g_I\) has engineering dimension \(D-d_I\), define the dimensionless
coordinate
\[
  \lambda_I(\mu)
  =
  \mu^{d_I-D}g_I(\mu).
\]
When the coordinate list contains all dimensionless ratios relevant to the
chosen family of theories, the equations may be written as
\[
  \frac{\dd\lambda_I}{\dd\log\mu}
  =
  \beta_I(\{\lambda_J\}).
\]
This autonomous form is a coordinate statement: the beta functions depend on
the selected operator basis, subtraction configurations, field normalization,
and finite part conventions.

The practical value of the 1PI RG is logarithmic organization.  A fixed-order
comparison between two widely separated scales can contain powers of
\(\log(\mu/\mu')\).  The differential equation breaks that comparison into
small steps and then integrates the result, thereby summing the logarithmic
terms determined by the computed beta function.

\section{Callan--Symanzik Equation from Bare-Scale Independence}

The beta function is a vector field on the renormalized coordinate space.  Its
action on correlation functions follows from the same construction of
renormalized coordinates.

There are two logically distinct field factors in the preceding discussion,
after the reference renormalization convention has been chosen.  The factor
\(Z_{\rm MOM}(\mu)\) is finite after renormalization; it relates an already
renormalized reference field \(\phi_R^{\rm ref}\) to the momentum-subtraction
field used in the scale-\(\mu\) coordinate chart,
\[
  [\phi]_\mu=Z_{\rm MOM}(\mu)^{-1/2}\phi_R^{\rm ref}.
\]
The factor denoted \(Z_\phi\) in the Callan--Symanzik equation is instead the
cutoff-dependent total factor that relates the regulated bare field to the
renormalized field chosen for the chart.  If
\[
  \phi_R^{\rm ref}
  =
  \bigl(Z_\phi^{\rm ref}(\lambda,\mu_{\rm ref},\Lambda)\bigr)^{-1/2}\phi_0 ,
\]
then the momentum-subtraction chart whose elementary field is \([\phi]_\mu\)
has
\[
  [\phi]_\mu
  =
  \bigl(Z_\phi^{\rm ref}(\lambda,\mu_{\rm ref},\Lambda)\,
  Z_{\rm MOM}(\mu)\bigr)^{-1/2}
  \phi_0 .
\]
Thus, in that chart,
\[
  Z_\phi^{\rm chart}(\lambda(\mu),\mu,\Lambda)
  =
  Z_\phi^{\rm ref}(\lambda,\mu_{\rm ref},\Lambda)\,Z_{\rm MOM}(\mu).
\]
Below the superscript ``chart'' is suppressed: \(Z_\phi\) always denotes the
total bare-to-renormalized field factor for the correlation functions being
studied, while \(Z_{\rm MOM}(\mu)\) denotes the finite momentum-subtraction rescaling
appearing in the 1PI vertex coordinates.

Let \(\phi_0\) be the regulated bare scalar field.  Let \(\phi_R\) now denote
the elementary field of the selected chart, for instance \([\phi]_\mu\) in the
momentum-subtraction chart above, and choose \(Z_\phi\) by
\[
  \phi_R=Z_\phi(\lambda(\mu),\mu,\Lambda)^{-1/2}\phi_0 .
\]
For a connected noncoincident \(n\)-point function, write
\[
  G_{0,\Lambda}^{(n)}(x_1,\ldots,x_n;g_0,\Lambda)
  =
  \langle
    \phi_0(x_1)\cdots\phi_0(x_n)
  \rangle_{\!{\rm conn},\Lambda,g_0}
\]
for the regulated bare function.  A renormalization chart consists of
functions
\[
  g_0^a=F^a(\lambda(\mu),\mu,\Lambda)
\]
and a field factor \(Z_\phi\) such that
\[
  G_R^{(n)}(x_1,\ldots,x_n;\lambda,\mu)
  =
  \lim_{\Lambda\to\infty}
  Z_\phi^{-n/2}
  G_{0,\Lambda}^{(n)}
  \bigl(x_1,\ldots,x_n;F(\lambda,\mu,\Lambda),\Lambda\bigr)
\]
exists as a distribution away from coincident points.  The finite coordinates
\(\lambda_I\) are dimensionless coordinates; dimensionful parameters are
either included as additional dimensionless ratios or are displayed separately
below.

Let \(t=\log\mu\).  The RG vector field in this chart is
\[
  \beta_I(\lambda)
  =
  \left.
  \frac{\dd\lambda_I}{\dd t}
  \right|_{g_0,\Lambda}.
\]
The anomalous dimension of the elementary field in the same chart is
\[
  \gamma_\phi(\lambda)
  =
  \frac12
  \left.
  \frac{\dd\log Z_\phi}{\dd t}
  \right|_{g_0,\Lambda}.
\]
The convention agrees with the normalization used earlier:
\(\phi_R=Z_\phi^{-1/2}\phi_0\).

\begin{proposition}[Callan--Symanzik equation]
With the definitions above, the renormalized connected noncoincident
\(n\)-point function satisfies
\begin{equation}
  \left(
    \mu\frac{\partial}{\partial\mu}
    +
    \sum_I\beta_I(\lambda)\frac{\partial}{\partial\lambda_I}
    +
    n\gamma_\phi(\lambda)
  \right)
  G_R^{(n)}
  =
  0 .
  \label{eq:callan-symanzik-connected}
\end{equation}
The derivative \(\mu\partial_\mu\) is taken at fixed \(x_i\) and fixed
renormalized coordinates \(\lambda_I\).
\end{proposition}

\begin{proof}
The regulated bare function contains no independent dependence on the
auxiliary scale \(\mu\).  Therefore
\[
  0
  =
  \left.
  \frac{\dd}{\dd t}
  G_{0,\Lambda}^{(n)}(x_1,\ldots,x_n;g_0,\Lambda)
  \right|_{g_0,\Lambda}.
\]
Using
\[
  G_{0,\Lambda}^{(n)}
  =
  Z_\phi^{n/2}G_R^{(n)}
\]
inside the renormalization chart and differentiating at fixed bare data gives
\[
  0
  =
  Z_\phi^{n/2}
  \left[
    \left(
      \frac{\partial}{\partial t}
      +
      \sum_I\beta_I\frac{\partial}{\partial\lambda_I}
    \right)G_R^{(n)}
    +
    \frac n2
    \frac{\dd\log Z_\phi}{\dd t}
    G_R^{(n)}
  \right].
\]
The cutoff limit exists by the definition of the chart, and the displayed
identity becomes \eqref{eq:callan-symanzik-connected}.
\end{proof}

Contact terms are not part of the displayed noncoincident equation.  When
insertion points collide, the omitted terms are local distributions supported
on the collision diagonals.  Their organization requires a renormalized local
operator basis and is developed in the next chapter.

\section{Canonical Scaling and Fixed Points}

The Callan--Symanzik equation records dependence on the auxiliary
renormalization scale.  Dimensional homogeneity is a separate statement.
Suppose the elementary scalar has engineering dimension
\[
  d_\phi^{\rm eng}=\frac{D-2}{2}
\]
and let \(u_A\) denote displayed dimensionful renormalized parameters with
engineering dimensions \(y_A\).  Away from coincident points, dimensional
homogeneity gives
\[
  \left(
    \sum_{i=1}^n x_i^\nu\frac{\partial}{\partial x_i^\nu}
    -
    \mu\frac{\partial}{\partial\mu}
    -
    \sum_A y_Au_A\frac{\partial}{\partial u_A}
    +
    n d_\phi^{\rm eng}
  \right)
  G_R^{(n)}
  =
  0 .
\]
If the \(u_A\) are displayed rather than converted into dimensionless ratios,
the Callan--Symanzik equation includes the additional term
\(\sum_A\beta_{u_A}\partial/\partial u_A\), where
\[
  \beta_{u_A}
  =
  \left.
  \frac{\dd u_A}{\dd\log\mu}
  \right|_{g_0,\Lambda}.
\]
Combining that equation with dimensional homogeneity gives
\[
  \left(
    \sum_{i=1}^n x_i^\nu\frac{\partial}{\partial x_i^\nu}
    +
    \sum_I\beta_I\frac{\partial}{\partial\lambda_I}
    +
    \sum_A(\beta_{u_A}-y_Au_A)\frac{\partial}{\partial u_A}
    +
    n(d_\phi^{\rm eng}+\gamma_\phi)
  \right)
  G_R^{(n)}
  =
  0 .
\]
Equivalently, one may replace \(u_A\) by dimensionless ratios
\(u_A/\mu^{y_A}\); then the canonical terms are included in the beta functions
of those ratios.

At a fixed point with no displayed dimensionful deformation,
\(\beta_I(\lambda_\ast)=0\) and
\(\gamma_\phi(\lambda_\ast)=\gamma_{\phi,\ast}\).  The noncoincident
correlator then obeys
\[
  \left(
    \sum_{i=1}^n x_i^\nu\frac{\partial}{\partial x_i^\nu}
    +
    n\Delta_\phi
  \right)
  G_R^{(n)}(x_1,\ldots,x_n;\lambda_\ast)
  =
  0,
  \qquad
  \Delta_\phi=d_\phi^{\rm eng}+\gamma_{\phi,\ast}.
\]
Thus the field scaling dimension at the fixed point is the engineering
dimension plus the anomalous dimension defined by the field normalization in
the renormalization chart.  Scaling dimensions of composite operators require
the mixing analysis of local operator insertions.

\section{Logarithmic Consistency}

The existence of finite beta functions imposes relations among logarithms in
cutoff perturbation theory.  Let
\[
  L=\log\frac{\Lambda}{\mu},
\]
and suppose a one-coupling coordinate has the bare expansion
\[
  g(\mu)
  =
  g_0-b_1g_0^2L-b_2g_0^3L-c_1g_0^3L^2+O(g_0^4).
\]
Since \(\dd L/\dd\log\mu=-1\),
\[
  \frac{\dd g(\mu)}{\dd\log\mu}
  =
  b_1g_0^2+b_2g_0^3+2c_1g_0^3L+O(g_0^4).
\]
Inverting the first displayed relation to the order needed gives
\[
  g_0=g(\mu)+b_1g(\mu)^2L+O(g(\mu)^3).
\]
Therefore
\[
  \frac{\dd g(\mu)}{\dd\log\mu}
  =
  b_1g(\mu)^2+b_2g(\mu)^3
  +
  2\bigl(c_1+b_1^2\bigr)g(\mu)^3L
  +
  O(g(\mu)^4).
\]
A beta function expressed in terms of the renormalized coordinate must be
finite as the cutoff is removed.  Hence the explicit \(L\)-term at this order
vanishes:
\[
  c_1=-b_1^2.
\]

Figure~\ref{fig:volume-ii-double-log-consistency} shows why the nested
one-loop double logarithm is constrained by the one-loop beta function: the
explicit cutoff logarithm cancels only for \(c_1=-b_1^2\).

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}[shift={(-4.2,0.15)}]
      \node[font=\small] at (0,2.0) {single logarithm};
      \draw[thick] (-2.0,0)--(-0.75,0);
      \draw[thick] (0.75,0)--(2.0,0);
      \draw[thick] (-0.75,0)..controls (-0.35,0.62) and (0.35,0.62)..(0.75,0);
      \draw[thick] (-0.75,0)..controls (-0.35,-0.62) and (0.35,-0.62)..(0.75,0);
      \node[blue!70!black] at (0,-1.05) {\(\sim b_1 g_0^2 L\)};
    \end{scope}

    \node[font=\large] at (-1.1,0.15) {\(\Longrightarrow\)};

    \begin{scope}[shift={(2.0,0.55)}]
      \node[font=\small] at (0,1.85) {nested four-point logarithms};
      \draw[thick] (-2.15,0)--(-0.9,0);
      \draw[thick] (0.9,0)--(2.15,0);
      \draw[thick] (-0.9,0)..controls (-0.45,0.72) and (0.45,0.72)..(0.9,0);
      \draw[thick] (-0.9,0)..controls (-0.45,-0.72) and (0.45,-0.72)..(0.9,0);
      \draw[orange!85!black, thick] (-0.45,0.35)..controls (-0.15,0.72) and
        (0.15,0.72)..(0.45,0.35);
      \draw[orange!85!black, thick] (-0.45,0.35)..controls (-0.15,0.02) and
        (0.15,0.02)..(0.45,0.35);
      \node[blue!70!black] at (0,-1.0) {\(\sim c_1g_0^3L^2\)};
    \end{scope}

    \node[font=\large] at (4.55,0.55) {\(+\)};

    \begin{scope}[shift={(6.55,0.55)}]
      \node[font=\small] at (0,1.85) {permuted channels};
      \coordinate (a) at (-0.9,0);
      \coordinate (b) at (0.9,0);
      \draw[thick] (-1.75,0.75)--(a)--(-1.75,-0.75);
      \draw[thick] (1.75,0.75)--(b)--(1.75,-0.75);
      \draw[thick] (a)..controls (-0.28,0.72) and (0.28,0.72)..(b);
      \draw[thick] (a)..controls (-0.28,-0.72) and (0.28,-0.72)..(b);
      \draw[orange!85!black, thick] (0,0.72)
        ..controls (0.45,1.25) and (-0.45,1.25)..(0,0.72);
      \fill (a) circle (1.8pt);
      \fill (b) circle (1.8pt);
      \node[blue!70!black] at (0,-1.0) {plus subleading single logs};
    \end{scope}

    \node[align=center, red!75!black] at (2.0,-1.75)
      {finite beta function\\requires\\\(c_1=-b_1^2\)};
  \end{tikzpicture}
  }
  \caption{The double logarithm associated with nested one-loop subgraphs is
  constrained by the one-loop beta function.  Its coefficient is fixed by the
  requirement that the beta function, expressed in the renormalized coordinate,
  have a finite cutoff-removal limit.}
  \label{fig:volume-ii-double-log-consistency}
\end{figure}

This calculation is the scale-differential form of the locality statements
from the previous chapters.  The logarithm associated with a subgraph can be
reused inside a larger graph, and the coefficient of the resulting double
logarithm is determined by the lower-order beta function rather than by an
independent datum.

\section{Scheme Changes}

A renormalization scheme is a coordinate system on the renormalized family of
theories.  Let \(\lambda\) and \(\widetilde\lambda\) be two coordinates related
by a finite redefinition
\[
  \widetilde\lambda=f(\lambda),
  \qquad
  f(0)=0,\qquad f'(0)=1,
\]
where the last condition fixes the leading normalization of the coordinate.
If
\[
  \frac{\dd\lambda}{\dd\log\mu}=\beta(\lambda),
  \qquad
  \frac{\dd\widetilde\lambda}{\dd\log\mu}
  =
  \widetilde\beta(\widetilde\lambda),
\]
then the chain rule gives
\[
  \widetilde\beta(f(\lambda))=f'(\lambda)\beta(\lambda),
  \qquad
  \beta(\lambda)=\frac{\widetilde\beta(f(\lambda))}{f'(\lambda)}.
\]
If only the first nonlinear coefficient is displayed,
\[
  f(\lambda)=\lambda+\alpha\lambda^2+O(\lambda^3),
\]
then, to first order in \(\alpha\) and without fixing the omitted cubic
coefficient,
\[
  \beta(\lambda)
  =
  \widetilde\beta(\lambda)
  +
  \alpha\lambda
  \left(
    \lambda\frac{\partial}{\partial\lambda}-2
  \right)
  \widetilde\beta(\lambda)
  +
  O(\alpha^2).
\]
For one classically marginal coordinate,
\[
  \beta(\lambda)=b_1\lambda^2+b_2\lambda^3+b_3\lambda^4+\cdots,
  \qquad
  \widetilde\beta(\lambda)
  =
  \widetilde b_1\lambda^2+\widetilde b_2\lambda^3
  +\widetilde b_3\lambda^4+\cdots .
\]
The relation above already implies
\[
  b_1=\widetilde b_1,
  \qquad
  b_2=\widetilde b_2 .
\]
To see the first scheme-dependent coefficient without hiding the role of the
next finite redefinition, write
\[
  \widetilde\lambda
  =
  \lambda+\alpha\lambda^2+\rho\lambda^3+O(\lambda^4).
\]
Using
\[
  \widetilde\beta(\widetilde\lambda)
  =
  f'(\lambda)\,\beta(\lambda),
  \qquad
  \widetilde\lambda=f(\lambda),
\]
and re-expanding in powers of \(\widetilde\lambda\) gives
\[
  \widetilde b_1=b_1,\qquad
  \widetilde b_2=b_2,\qquad
  \widetilde b_3
  =
  b_3-\alpha b_2+(\rho-\alpha^2)b_1 .
\]
Thus, for a single marginal coupling and with the leading normalization fixed,
the first two perturbative coefficients are invariant under finite analytic
reparametrizations, while later coefficients are coordinate dependent.

For several couplings the corresponding statement is a statement about jets of
a vector field, not about a list of scalar coefficients.  Let a finite
renormalized coordinate chart have coordinates \(\lambda^I\) centered at the
point under discussion, and expand
\begin{equation}
  \beta^I(\lambda)
  =
  M^I{}_J\lambda^J
  +
  \frac12 B^I{}_{JK}\lambda^J\lambda^K
  +
  O(\lambda^3),
  \qquad
  B^I{}_{JK}=B^I{}_{KJ}.
  \label{eq:multicoupling-beta-jet}
\end{equation}
Let another scheme be related by the finite analytic coordinate change
\begin{equation}
  \widetilde\lambda^A
  =
  R^A{}_I\lambda^I
  +
  \frac12 C^A{}_{IJ}\lambda^I\lambda^J
  +
  O(\lambda^3),
  \qquad
  C^A{}_{IJ}=C^A{}_{JI},
  \label{eq:multicoupling-coordinate-jet}
\end{equation}
where \(R\) is invertible.  Write \(S=R^{-1}\), and write
\(\widetilde M=RMS\).  Then the quadratic coefficient of
\(\widetilde\beta\) is
\begin{equation}
  \begin{aligned}
  \widetilde B^A{}_{BC}
  &=
  R^A{}_I B^I{}_{JK}S^J{}_B S^K{}_C  \\
  &\quad
  + C^A{}_{IJ}
    \bigl(S^I{}_B M^J{}_K S^K{}_C
          +S^I{}_C M^J{}_K S^K{}_B\bigr)
  - \widetilde M^A{}_D C^D{}_{IJ}S^I{}_B S^J{}_C .
  \end{aligned}
  \label{eq:multicoupling-quadratic-beta-transform}
\end{equation}
Thus the quadratic beta-function jet is not, in general, a tensorial invariant
of the renormalization scheme.  It is changed by the second derivative of the
coordinate map whenever the linear term \(M\) is present.  The invariant
linear data at a fixed point are the similarity class of \(M\), and hence its
eigenvalues.

\begin{proof}
The beta function transforms as a vector field:
\[
  \widetilde\beta^A(\widetilde\lambda)
  =
  \frac{\partial\widetilde\lambda^A}{\partial\lambda^I}
  \,\beta^I(\lambda).
\]
The inverse coordinate map has the expansion
\[
  \lambda^I
  =
  S^I{}_A\widetilde\lambda^A
  -
  \frac12 S^I{}_A C^A{}_{JK}S^J{}_B S^K{}_C
  \widetilde\lambda^B\widetilde\lambda^C
  +
  O(\widetilde\lambda^3).
\]
Substituting this inverse expansion and
\[
  \frac{\partial\widetilde\lambda^A}{\partial\lambda^I}
  =
  R^A{}_I+C^A{}_{IJ}\lambda^J+O(\lambda^2)
\]
into the vector-field transformation law and collecting the terms quadratic in
\(\widetilde\lambda\) gives
\eqref{eq:multicoupling-quadratic-beta-transform}.
\end{proof}

There is an important special case.  If all displayed couplings are
classically marginal at the reference point and the coordinate origin is the
Gaussian fixed point, then \(M=0\).  In that case
\eqref{eq:multicoupling-quadratic-beta-transform} reduces to
\[
  \widetilde B^A{}_{BC}
  =
  R^A{}_I B^I{}_{JK}S^J{}_B S^K{}_C .
\]
After the leading linear normalization \(R\) has been fixed, the one-loop
quadratic tensor \(B\) is scheme independent.  Without fixing \(R\), the same
geometric object is represented by different components in different linear
bases.  If \(M\neq0\), for example when relevant, irrelevant, or explicitly
dimensionful coordinates are included, the quadratic coefficients are already
coordinate dependent.  At cubic order, even in the purely marginal case, the
quadratic coordinate redefinition \(C\) combines with \(B\); hence the
single-coupling statement \(b_2=\widetilde b_2\) has no componentwise
multi-coupling analogue without additional structure.

\begin{theorem}[Scheme equivalence for matched physical observables]
\label{thm:scheme-equivalence-physical-observables}
Let \(S_{\Lambda_0}(\theta)\) be the local action used in the regulated path
integral, including all regulator-dependent local terms required by the chosen
regularization procedure.  Here \(\Lambda_0\) denotes the ultraviolet regulator
and \(\theta\) denotes the full list of regulated action parameters.  Let
\(R\) and \(\widetilde R\) be two renormalization schemes on the same
perturbative domain.  Assume the following data.
\begin{enumerate}
  \item Each scheme assigns a finite renormalized coordinate to the same
  regulated action parameter \(\theta\):
  \[
    \lambda_\Lambda=\lambda_R(\theta,\Lambda_0),
    \qquad
    \widetilde\lambda_\Lambda=
    \lambda_{\widetilde R}(\theta,\Lambda_0),
  \]
  and the cutoff-removal limits are related by a finite analytic matching map
  \[
    \widetilde\lambda=f(\lambda),
    \qquad f(0)=0,\qquad Df(0)\ \hbox{invertible}.
  \]
  The same statement includes the finite field normalizations and local source
  coordinate maps needed to define inserted operators.
  \item The observable \(A_\Lambda\) is defined at the regulated level before a
  renormalization scheme is chosen.  Examples are an on-shell transition
  probability between normalized Haag--Ruelle particle states, a perturbative
  LSZ matrix element after the pole positions and residues have been matched,
  or an infrared-safe inclusive observable with the measurement function held
  fixed.
  \item The ultraviolet and infrared limits required to define \(A\) exist
  order by order in the perturbative domain under the matching above.  In
  massless theories this includes the infrared prescription, such as
  nonexceptional Euclidean momenta for Green functions or an inclusive
  measurement for scattering.
\end{enumerate}
Let \(A_R(\lambda)\) and
\(A_{\widetilde R}(\widetilde\lambda)\) be the perturbative expansions obtained
by computing the same regulated observable in the two schemes and then taking
the common limit.  Then, as formal perturbative series,
\begin{equation}
  A_R(\lambda)
  =
  A_{\widetilde R}(f(\lambda)).
  \label{eq:scheme-equivalence-observable}
\end{equation}
\end{theorem}

\begin{proof}
At finite \(\Lambda_0\), \(A_\Lambda\) is a number, distribution, or transition
probability computed in the regulated theory with action parameter \(\theta\).  It
does not know which finite coordinates will later be used to label the same
regulated action.  Thus
\[
  A_\Lambda(\theta)
  =
  A_{\Lambda,R}(\lambda_R(\theta,\Lambda_0))
  =
  A_{\Lambda,\widetilde R}
    (\lambda_{\widetilde R}(\theta,\Lambda_0))
\]
after including the field and source normalizations that belong to the two
coordinate charts.  The matching map is defined by eliminating the regulated
action parameters between the two coordinate assignments and then taking
\(\Lambda_0\to\infty\):
\[
  \lambda_{\widetilde R}(\theta,\Lambda_0)
  =
  f_{\Lambda_0}(\lambda_R(\theta,\Lambda_0)),
  \qquad
  f_{\Lambda_0}\to f .
\]
If the observable has the stated perturbative limit, the finite-regulator
identity passes to the limit order by order and gives
\eqref{eq:scheme-equivalence-observable}.

For an LSZ amplitude, the finite field rescaling between schemes changes
off-shell Green functions and amputated kernels, but the residue and
one-particle-state normalizations used in the LSZ construction change by the
inverse factors.  The matched on-shell matrix element is therefore the same
observable.  For a local source insertion, a finite local source-coordinate
redefinition can change the coincident-point extension of the source
functional.  This does not change a separated-point observable.  If the
observable itself probes coincident support, then its definition includes that
source-response convention and the convention must be matched as part of the
observable data.
\end{proof}

\begin{remark}[What is not scheme invariant]
Equation~\eqref{eq:scheme-equivalence-observable} is a statement with a
matching map.  It does not say that
\(A_R(\lambda)=A_{\widetilde R}(\lambda)\) after the same coordinate values are
inserted in two different schemes.  It also does not make off-shell Green
functions, beta-function components, anomalous-dimension matrices, or contact
terms coordinate independent.  Those quantities are useful precisely because
they are chart-dependent representatives from which matched physical
observables, fixed-point eigenvalues, anomaly classes, and other invariant data
are extracted.
\end{remark}

\section{Perturbative Series as Asymptotic Data}

The beta functions, anomalous dimensions, fixed-point coordinates, and
critical exponents computed in the preceding and following chapters are
usually obtained as formal expansions.  The meaning of such an expansion must
be specified separately from the rules used to compute its coefficients.

\begin{definition}[Asymptotic expansion]
Let \(A(g)\) be defined for \(g\) in a sector \(S\subset\mathbb C\) whose
closure contains \(g=0\), and let \(\sum_{n\ge0}a_n g^n\) be a formal power
series.  The series is an asymptotic expansion of \(A\) in \(S\) if, for every
\(N\ge0\),
\[
  A(g)-\sum_{n=0}^{N}a_ng^n
  =
  o(g^N)
  \qquad (g\to0,\ g\in S')
\]
on every closed subsector \(S'\) strictly contained in \(S\).  A stronger
version replaces \(o(g^N)\) by \(O(g^{N+1})\) with constants depending on
\(N\) and \(S'\).
\end{definition}

An asymptotic expansion need not converge.  Even when it determines every
coefficient \(a_n\), it does not by itself determine the function \(A\):
functions differing by terms such as \(\exp(-c/g)\), with \(c>0\) in the
sector under consideration, have the same power-series asymptotic expansion at
the origin.  Such nonperturbative terms are invisible to all orders in
ordinary perturbation theory.

The common mechanism behind divergence is factorial growth of coefficients.
In finite-dimensional integrals and in many field-theoretic expansions, the
number of diagrams grows factorially.  In examples where the saddle analysis
of the analytically continued action is under control, the large-order
coefficients have the form
\[
  a_n\sim C\,(-A)^{-n}\,\Gamma(n+b)
\]
for constants \(A,C,b\) depending on the observable and on the expansion
coordinate.  Dyson's instability argument for QED is the simplest physical
warning: if a perturbative series in \(e^2\) had a nonzero radius of
convergence around \(e^2=0\), it would analytically continue to negative
\(e^2\), where like charges attract and the vacuum is not stable in the same
way.\footnote{See F.~J.~Dyson, ``Divergence of perturbation theory in quantum
electrodynamics,'' \emph{Phys. Rev.} \textbf{85} (1952) 631--632,
doi:10.1103/PhysRev.85.631.  For explicit large-order analysis in closely
related settings, see C.~M.~Bender and T.~T.~Wu, ``Anharmonic oscillator,''
\emph{Phys. Rev.} \textbf{184} (1969) 1231--1260, doi:10.1103/PhysRev.184.1231,
and L.~N.~Lipatov, ``Divergence of the perturbation-theory series and the
quasiclassical theory,'' \emph{Sov. Phys. JETP} \textbf{45} (1977) 216--223.}
This argument is not a theorem about every renormalized quantity in every QFT;
it is a reason not to identify perturbation theory with a convergent
definition unless convergence or summability has actually been proved.

One possible reconstruction mechanism is Borel--Laplace summation.  It is a
theorem about a formal series together with analytic continuation and growth
data for its Borel transform.  The hypotheses are part of the statement.

\begin{definition}[Gevrey-one series and directional Borel sum]
\label{def:gevrey-one-directional-borel-sum}
Let
\[
  \widehat A(g)=\sum_{n\ge0}a_n g^n
\]
be a formal power series.  It is \emph{Gevrey-one} if there are constants
\(C,M>0\) such that
\[
  |a_n|\le C\,M^n n!
  \qquad(n\ge0).
\]
Its formal Borel transform is
\[
  \mathcal B\widehat A(\xi)
  =
  \sum_{n\ge0}\frac{a_n}{n!}\xi^n .
  \label{eq:formal-borel-transform}
\]
Fix a direction \(\theta\in\mathbb R/2\pi\mathbb Z\).  The direction is
admissible for \(\widehat A\) if the series
\(\mathcal B\widehat A\) converges near \(\xi=0\), continues holomorphically
to a sector
\[
  \Sigma_{\theta,\eta}
  =
  \{\xi\ne0:\ |\arg\xi-\theta|<\eta\}
\]
for some \(\eta>0\), and obeys an exponential bound
\[
  |\mathcal B\widehat A(\xi)|
  \le C_\eta \exp(|\xi|/R_\eta)
  \qquad(\xi\in\Sigma_{\theta,\eta})
  \label{eq:borel-exponential-bound}
\]
for some \(C_\eta,R_\eta>0\).  For \(g\) satisfying
\[
  |\arg g-\theta|<\frac\pi2-\delta,
  \qquad
  |g|<R_\eta\cos\delta
\]
with \(0<\delta<\pi/2\), the directional Borel sum is
\[
  \mathcal S_\theta \widehat A(g)
  =
  \frac1g\int_0^{\ee^{i\theta}\infty}
  \ee^{-\xi/g}\,\mathcal B\widehat A(\xi)\,\dd\xi ,
  \label{eq:directional-borel-sum}
\]
where the ray is oriented from \(0\) to infinity.
\end{definition}

\begin{theorem}[Watson lemma for the Borel--Laplace transform]
\label{thm:watson-borel-laplace}
Under the hypotheses of Definition~\ref{def:gevrey-one-directional-borel-sum},
\(\mathcal S_\theta\widehat A(g)\) is holomorphic in the indicated sector of
the \(g\)-plane and has \(\widehat A(g)\) as its asymptotic expansion there:
for every \(N\ge0\) and every closed subsector strictly inside the domain,
\[
  \mathcal S_\theta\widehat A(g)
  -
  \sum_{n=0}^N a_n g^n
  =
  O(g^{N+1}).
  \label{eq:borel-watson-remainder}
\]
\end{theorem}

\begin{proof}
Rotate the integration variable so that the Borel ray is the positive real
axis; the general case is the same after replacing \(g\) by
\(\ee^{-i\theta}g\).  The exponential bound
\eqref{eq:borel-exponential-bound} gives absolute convergence of
\eqref{eq:directional-borel-sum} on each closed subsector with
\(\Re(1/g)>1/R_\eta+\epsilon\).  Holomorphy follows by dominated convergence
applied to compact subsets of that domain.

Choose \(\rho>0\) inside the convergence disk of the Borel transform.  For
\(|\xi|\le\rho\),
\[
  \mathcal B\widehat A(\xi)
  =
  \sum_{n=0}^N\frac{a_n}{n!}\xi^n+\xi^{N+1}R_N(\xi)
\]
with \(R_N\) bounded on \(|\xi|\le\rho\).  The polynomial terms give exactly
\[
  \frac1g\int_0^\infty \ee^{-\xi/g}\frac{\xi^n}{n!}\,\dd\xi
  =
  g^n
\]
when \(\Re g>0\), and by the same rotated-ray computation in a sector.  The
small-\(\xi\) remainder is bounded by a constant times
\[
  \frac1{|g|}\int_0^\rho
  \ee^{-c\xi/|g|}\xi^{N+1}\,\dd\xi
  =
  O(|g|^{N+1})
\]
on each closed subsector, where \(c>0\) is the lower bound for
\(\Re(\xi/g)/(|\xi|/|g|)\).  The tail integral from \(\rho\) to infinity is
bounded by \(C\exp(-c'\rho/|g|)\) for some \(c'>0\), hence is
\(O(|g|^{N+1})\) for every fixed \(N\).  This proves
\eqref{eq:borel-watson-remainder}.
\end{proof}

\begin{definition}[Lateral Borel sums and Stokes discontinuity]
\label{def:lateral-borel-sums-stokes-discontinuity}
Let \(\widehat A(g)\) be Gevrey-one, and suppose that its Borel transform is
holomorphic in a punctured sector around the ray
\(\arg\xi=\theta\), except for a locally finite set of singular points on the
ray.  If the deformed rays
\(\arg\xi=\theta\pm\epsilon\) have Borel--Laplace sums and these sums have
limits as \(\epsilon\downarrow0\), the limits are the lateral sums
\[
  \mathcal S_{\theta^+}\widehat A(g),
  \qquad
  \mathcal S_{\theta^-}\widehat A(g).
\]
Their difference
\[
  \operatorname{Disc}_\theta \widehat A(g)
  =
  \mathcal S_{\theta^+}\widehat A(g)
  -
  \mathcal S_{\theta^-}\widehat A(g)
  \label{eq:borel-stokes-discontinuity}
\]
is the Stokes discontinuity across the ray.  This definition is a statement
about boundary values of a named Borel transform; it is not a prescription
unless the analytic continuation, growth bounds, and lateral limits have been
established.
\end{definition}

\begin{proposition}[Simple-pole Stokes jump]
\label{prop:simple-pole-stokes-jump}
Let \(A>0\), and suppose that in a neighborhood of \(\xi=A\) the Borel
transform has the form
\[
  \mathcal B\widehat A(\xi)
  =
  \frac{r}{\xi-A}+H(\xi),
  \label{eq:simple-borel-pole-local-form}
\]
where \(H\) is holomorphic.  Assume that away from this singularity the
lateral Borel integrals along the positive ray satisfy the hypotheses of
Definition~\ref{def:lateral-borel-sums-stokes-discontinuity}.  Then the
contribution of this pole to the upper-minus-lower discontinuity is
\[
  \operatorname{Disc}_{0}\widehat A(g)
  =
  \frac{2\pi\ii r}{g}\,\ee^{-A/g}
  +
  \operatorname{Disc}_{0}^{\rm other}\widehat A(g),
  \label{eq:simple-pole-stokes-jump}
\]
where the last term denotes contributions from other singularities on the
same ray.
\end{proposition}

\begin{proof}
Subtract the holomorphic part \(H\), whose upper and lower deformations across
a small disk around \(A\) have the same limit.  The difference between the
two lateral contours for the pole term is a positively oriented small circle
around \(A\).  Hence
\[
  \frac1g\oint_{|\xi-A|=\epsilon}
  \ee^{-\xi/g}\frac{r}{\xi-A}\,\dd\xi
  =
  \frac{2\pi\ii r}{g}\,\ee^{-A/g}.
\]
The remaining parts of the contour cancel by deformation in the common
holomorphic domain.  Summing over singularities on the ray gives the stated
local contribution.
\end{proof}

\begin{definition}[One-parameter transseries datum]
\label{def:one-parameter-transseries-datum}
Fix \(A>0\), \(\beta\in\mathbb C\), and a sector in the \(g\)-plane.  A
one-parameter transseries datum at the first exponential scale is a formal
object
\[
  \widehat F(g;\sigma)
  =
  \widehat\Phi_0(g)
  +
  \sigma\,\ee^{-A/g}g^\beta\widehat\Phi_1(g),
  \label{eq:one-parameter-transseries-datum}
\]
where \(\widehat\Phi_0\) and \(\widehat\Phi_1\) are formal power series and
\(\sigma\in\mathbb C\) is a parameter.  A summation of
\eqref{eq:one-parameter-transseries-datum} requires summation data for both
formal series and a rule for how \(\sigma\) changes when the summation
direction crosses a Stokes ray.
\end{definition}

\begin{proposition}[Algebra of the first Stokes cancellation]
\label{prop:first-stokes-cancellation-algebra}
Assume that the lateral sums of \(\widehat\Phi_0\) and \(\widehat\Phi_1\)
exist on a ray and satisfy
\[
  \mathcal S_{+}\widehat\Phi_0
  -
  \mathcal S_{-}\widehat\Phi_0
  =
  C\,\ee^{-A/g}g^\beta\,\mathcal S_{-}\widehat\Phi_1,
  \qquad
  \mathcal S_{+}\widehat\Phi_1=\mathcal S_{-}\widehat\Phi_1,
  \label{eq:first-stokes-cancellation-hypothesis}
\]
for a constant \(C\).  Then the two lateral sums of
\(\widehat F(g;\sigma)\) agree if and only if
\[
  \sigma_-=\sigma_+ + C .
  \label{eq:transseries-parameter-stokes-jump}
\]
If the data are real on the two sides of a real Stokes ray and \(C\) is purely
imaginary, the median choice
\[
  \sigma_+=\sigma_{\rm R}-\frac{C}{2},
  \qquad
  \sigma_-=\sigma_{\rm R}+\frac{C}{2},
  \qquad
  \sigma_{\rm R}\in\mathbb R,
  \label{eq:median-transseries-parameter}
\]
gives a common real value whenever the two lateral boundary values are complex
conjugates.
\end{proposition}

\begin{proof}
Using \eqref{eq:first-stokes-cancellation-hypothesis},
\[
\begin{aligned}
  \mathcal S_{+}\widehat F(g;\sigma_+)
  &=
  \mathcal S_{-}\widehat\Phi_0
  +(C+\sigma_+)\ee^{-A/g}g^\beta\mathcal S_{-}\widehat\Phi_1,\\
  \mathcal S_{-}\widehat F(g;\sigma_-)
  &=
  \mathcal S_{-}\widehat\Phi_0
  +\sigma_-\ee^{-A/g}g^\beta\mathcal S_{-}\widehat\Phi_1.
\end{aligned}
\]
Equality is therefore equivalent to \eqref{eq:transseries-parameter-stokes-jump}
unless the displayed one-instanton lateral sum vanishes identically, in which
case the same equation is the natural continuous extension.  The median
choice is the symmetric solution of this affine relation.  Under the stated
reality hypothesis the two lateral values are conjugate, so their common
symmetric value is real.
\end{proof}

\begin{example}[Zero-dimensional stable quartic integral]
\label{ex:zero-dimensional-quartic-large-order}
The normalized integral
\[
  I(g)
  =
  \frac1{\sqrt{2\pi}}
  \int_{\mathbb R}
  \exp\left(-\frac{x^2}{2}-\frac{g x^4}{4!}\right)\,\dd x
  \qquad(\Re g>0)
  \label{eq:zero-dimensional-quartic-integral}
\]
is a finite-dimensional model of the distinction between coefficientwise
perturbation theory and analytic reconstruction.  Termwise expansion at
\(g=0\) gives
\[
  I(g)\sim\sum_{n\ge0}a_n g^n,
  \qquad
  a_n
  =
  (-1)^n\frac{(4n)!}{n!\,96^n(2n)!}.
  \label{eq:zero-dimensional-quartic-coefficients}
\]
Indeed
\[
  \frac1{\sqrt{2\pi}}\int_{\mathbb R}x^{4n}\ee^{-x^2/2}\,\dd x
  =
  (4n-1)!!
  =
  \frac{(4n)!}{2^{2n}(2n)!},
\]
and the extra factor is \((-1)^n/(n!(4!)^n)\).  Stirling's formula gives
\[
  \frac{a_{n+1}}{(n+1)a_n}\longrightarrow -\frac23 ,
  \label{eq:zero-dimensional-quartic-ratio}
\]
so the perturbative series has zero radius of convergence.  Dividing by
\(n!\) gives the Borel series
\[
  \mathcal B I(\xi)
  =
  \sum_{n\ge0}\frac{a_n}{n!}\xi^n
  =
  {}_2F_1\left(\frac14,\frac34;1;-\frac{2\xi}{3}\right)
  \label{eq:zero-dimensional-quartic-borel-hypergeometric}
\]
near the origin.  The hypergeometric function has its first finite singularity
at argument \(1\), hence at \(\xi=-3/2\) in this normalization.  Along the
positive \(g\)-axis the integral is nevertheless well-defined and its
perturbative series is a candidate for ordinary Borel summation.  The
opposite sign of \(g\) makes the integral unstable on the real contour; the
same analytic formula must then be understood only after choosing a complex
cycle.
\end{example}

\begin{proposition}[Conformal Borel map for one negative cut]
\label{prop:conformal-borel-negative-cut-map}
Let \(A>0\), and suppose a Borel transform \(B(\xi)\) is holomorphic on the
cut plane \(\mathbb C\setminus(-\infty,-A]\).  The map
\[
  w(\xi)
  =
  \frac{\sqrt{1+\xi/A}-1}{\sqrt{1+\xi/A}+1},
  \qquad
  \xi(w)=\frac{4Aw}{(1-w)^2},
  \label{eq:conformal-borel-negative-cut-map}
\]
with the square root chosen positive at \(\xi=0\), maps the cut plane
biholomorphically to the unit disk.  It sends \(\xi=0\) to \(w=0\), the first
branch point \(\xi=-A\) to \(w=-1\), and the positive real \(\xi\)-axis to
\(0\le w<1\).
\end{proposition}

\begin{proof}
Set \(y=\sqrt{1+\xi/A}\).  The cut plane is mapped by \(y\) to the right half
plane \(\Re y>0\).  The Cayley transform \(w=(y-1)/(y+1)\) maps the right
half plane to \(|w|<1\).  Solving \(y=(1+w)/(1-w)\) gives
\[
  1+\frac{\xi}{A}=\left(\frac{1+w}{1-w}\right)^2,
  \qquad
  \xi=\frac{4Aw}{(1-w)^2}.
\]
The stated images of \(0\), \(-A\), and the positive real ray are immediate.
\end{proof}

\begin{remark}[Status of conformal-Borel numerics]
\label{rem:conformal-borel-numerics-status}
The conformal map in
Proposition~\ref{prop:conformal-borel-negative-cut-map} is a rigorous
acceleration device only when the analytic domain of the Borel transform is
known.  In many applications to critical exponents one assumes, from
finite-dimensional saddle analysis or from additional field-theoretic input,
that the nearest Borel singularity is on the negative real axis and that no
positive-ray singularity obstructs the Laplace integral.  A conformal-Borel
or Borel--Pade computation made under those assumptions is a controlled
numerical procedure conditional on the stated singularity hypothesis; it is
not a proof that the regulated QFT exists or that the perturbative expansion
by itself defines it.
\end{remark}

\begin{definition}[Borel--Leroy and conformal-Borel approximants]
\label{def:borel-leroy-conformal-approximants}
Let
\(\widehat A(g)=\sum_{n\ge0}a_n g^n\) be a formal series and let \(b>-1\).
The Borel--Leroy transform with parameter \(b\) is
\[
  \mathcal B_b\widehat A(\xi)
  =
  \sum_{n\ge0}\frac{a_n}{\Gamma(n+b+1)}\xi^n .
  \label{eq:borel-leroy-transform}
\]
If \(\mathcal B_b\widehat A\) satisfies the same analytic-continuation and
growth hypotheses as in Definition~\ref{def:gevrey-one-directional-borel-sum},
with the branch of \(\xi^b\) fixed on the chosen ray, the corresponding
Borel--Leroy sum is
\[
  \mathcal S_{\theta,b}\widehat A(g)
  =
  \frac1{g^{b+1}}
  \int_0^{\ee^{i\theta}\infty}
  \ee^{-\xi/g}\xi^b
  \mathcal B_b\widehat A(\xi)\,\dd\xi .
  \label{eq:borel-leroy-sum}
\]
For the one-cut hypothesis of
Proposition~\ref{prop:conformal-borel-negative-cut-map}, write
\[
  \mathcal B_b\widehat A(\xi(w))
  =
  \sum_{m\ge0}c_m w^m
\]
near \(w=0\), and let
\[
  \mathcal B_{b,N}^{\rm conf}(\xi)
  =
  \sum_{m=0}^N c_m w(\xi)^m .
  \label{eq:truncated-conformal-borel-transform}
\]
The \(N\)-th conformal-Borel--Leroy approximant on the positive ray is
\[
  \mathcal S_{N}^{(A,b)}\widehat A(g)
  =
  \frac1{g^{b+1}}
  \int_0^\infty
  \ee^{-\xi/g}\xi^b
  \mathcal B_{b,N}^{\rm conf}(\xi)\,\dd\xi ,
  \qquad g>0,
  \label{eq:conformal-borel-leroy-approximant}
\]
when the integral converges.  The data \((A,b,N)\) are part of the
approximation, not hidden conventions.
\end{definition}

\begin{proposition}[Coefficient preservation by the truncated conformal map]
\label{prop:conformal-borel-leroy-coefficient-preservation}
Assume the notation of
Definition~\ref{def:borel-leroy-conformal-approximants}.  Suppose
\(\mathcal B_b\widehat A\) is holomorphic at the origin, and \(w(\xi)\) is the
map of Proposition~\ref{prop:conformal-borel-negative-cut-map}.  Then
\[
  \mathcal B_{b,N}^{\rm conf}(\xi)
  -
  \mathcal B_b\widehat A(\xi)
  =
  O(\xi^{N+1})
  \qquad(\xi\to0).
  \label{eq:conformal-borel-truncation-order}
\]
Consequently the formal small-\(g\) expansion of
\(\mathcal S_N^{(A,b)}\widehat A(g)\) reproduces the coefficients
\(a_0,\ldots,a_N\):
\[
  \mathcal S_N^{(A,b)}\widehat A(g)
  \sim
  \sum_{n=0}^N a_n g^n
  +
  O(g^{N+1})
  \label{eq:conformal-borel-leroy-reproduces-input}
\]
in the same Watson-lemma sense as
Theorem~\ref{thm:watson-borel-laplace}, provided the positive-ray integral has
a uniform exponential-growth bound.
\end{proposition}

\begin{proof}
The map \(w(\xi)\) is holomorphic near \(\xi=0\), satisfies \(w(0)=0\), and
has nonzero derivative there:
\[
  w(\xi)=\frac{\xi}{4A}+O(\xi^2).
\]
It therefore defines an invertible holomorphic coordinate near the origin.
Truncating the Taylor expansion of
\(\mathcal B_b\widehat A(\xi(w))\) at order \(N\) in \(w\) changes the function
by \(O(w^{N+1})\), hence by \(O(\xi^{N+1})\).  This proves
\eqref{eq:conformal-borel-truncation-order}.

For each \(0\le n\le N\), the coefficient of \(\xi^n\) in
\(\mathcal B_{b,N}^{\rm conf}\) is therefore
\(a_n/\Gamma(n+b+1)\).  Termwise integration of the polynomial part gives
\[
  \frac1{g^{b+1}}
  \int_0^\infty
  \ee^{-\xi/g}\xi^{n+b}\,\dd\xi
  =
  g^n\Gamma(n+b+1),
  \qquad b>-1,
  \label{eq:borel-leroy-gamma-normalization}
\]
and hence recovers \(a_n g^n\).  The remainder estimate is the same small-ray
and tail estimate used in
Theorem~\ref{thm:watson-borel-laplace}, with \(\xi^{N+1+b}\) replacing
\(\xi^{N+1}\).  The hypothesis \(b>-1\) is exactly the integrability condition
at \(\xi=0\).
\end{proof}

\begin{remark}[How this applies to critical exponents]
\label{rem:critical-exponent-borel-practice-status}
In the Wilson--Fisher \(\varepsilon\)-expansion and in fixed-dimension
renormalized perturbation theory, one often applies
\eqref{eq:conformal-borel-leroy-approximant} to a critical exponent, an
anomalous dimension, or a fixed-point coordinate.  The theorem-level input is
the list of loop coefficients together with the renormalized chart, the
Borel--Leroy parameter \(b\), the assumed first
singularity distance \(A\), the analytic domain in the Borel plane, and a rule
for estimating the truncation and parameter sensitivity.  When these analytic
data are not proved, the resulting number is a controlled extrapolation under
declared hypotheses; it is not a proof of the existence or uniqueness of the
critical QFT.
\end{remark}

The preceding example is the finite-dimensional form of Dyson's warning:
analytic continuation in a coupling can change the convergence of the
integration cycle and the physical stability of the problem.  The correct
mathematical object is therefore not only a formal series but a pair consisting
of a formal expansion and a choice of analytic sector or integration cycle.
This is the point of the Lefschetz-thimble formulation.

\begin{definition}[Holomorphic oscillatory integral and thimbles]
\label{def:lefschetz-thimble-data}
Let \(X\) be a complex \(n\)-manifold with holomorphic volume form \(\Omega\),
let \(S:X\to\mathbb C\) be a holomorphic Morse function, and let
\(\Gamma\) be a real \(n\)-cycle on which
\[
  I_\Gamma(g)=\int_\Gamma \ee^{-S(z)/g}a(z)\,\Omega
  \label{eq:holomorphic-oscillatory-integral}
\]
converges for \(g=|g|\ee^{i\theta}\).  Here \(a\) is holomorphic, or more
generally has controlled analytic growth compatible with the convergence of
the integral.  Set
\[
  h_\theta(z)=\operatorname{Re}\bigl(\ee^{-i\theta}S(z)\bigr),
  \qquad
  \varphi_\theta(z)=\operatorname{Im}\bigl(\ee^{-i\theta}S(z)\bigr).
\]
For a critical point \(p_\sigma\), its upward thimble
\(\mathcal J_\sigma^\theta\) is the real \(n\)-manifold obtained by flowing
out of \(p_\sigma\) along the gradient flow of \(h_\theta\), so that
\(h_\theta\to+\infty\) at the ends.  Its dual downward thimble
\(\mathcal K_\sigma^\theta\) is defined with the opposite flow.  Along either
flow,
\[
  \frac{\dd h_\theta}{\dd t}
  =
  \|\partial(\ee^{-i\theta}S)\|^2,
  \qquad
  \frac{\dd\varphi_\theta}{\dd t}=0
  \label{eq:thimble-flow-monotonicity}
\]
for the upward convention.  Thus the phase of the exponential is constant on
each thimble.
\end{definition}

\begin{theorem}[Finite-dimensional thimble decomposition]
\label{thm:finite-dimensional-thimble-decomposition}
Assume that \(h_\theta\) is proper on the relative cycle problem defined by
\eqref{eq:holomorphic-oscillatory-integral}, that all critical values
\(\ee^{-i\theta}S(p_\sigma)\) have distinct imaginary parts, and that the
flows have no escape at finite flow time.  Then the relative homology class of
\(\Gamma\) decomposes as
\[
  [\Gamma]
  =
  \sum_\sigma n_\sigma(\theta)
  [\mathcal J_\sigma^\theta],
  \qquad
  n_\sigma(\theta)=\langle \Gamma,\mathcal K_\sigma^\theta\rangle\in\mathbb Z,
  \label{eq:thimble-homology-decomposition}
\]
and
\[
  I_\Gamma(g)
  =
  \sum_\sigma n_\sigma(\theta)
  I_\sigma(g),
  \qquad
  I_\sigma(g)=
  \int_{\mathcal J_\sigma^\theta}
  \ee^{-S(z)/g}a(z)\,\Omega .
  \label{eq:thimble-integral-decomposition}
\]
Near \(g=0\), each \(I_\sigma(g)\) has a saddle expansion
\[
  I_\sigma(g)
  \sim
  \ee^{-S(p_\sigma)/g}
  (2\pi g)^{n/2}
  \frac{a(p_\sigma)}
       {\sqrt{\det H_\sigma}}
  \left(1+\sum_{m\ge1}c_{\sigma,m}g^m\right),
  \label{eq:thimble-saddle-expansion}
\]
after choosing the square-root phase determined by the oriented thimble; here
\(H_\sigma\) is the Hessian of \(S\) at \(p_\sigma\) in holomorphic local
coordinates.  When two critical values have the same
\(\varphi_\theta\), the integer basis
\([\mathcal J_\sigma^\theta]\) can jump by a Stokes transformation.  The
homology class \([\Gamma]\) and the integral \(I_\Gamma(g)\) are unchanged;
only their expression in the thimble basis changes.
\end{theorem}

\begin{proof}
The gradient equation in Definition~\ref{def:lefschetz-thimble-data} gives
\[
  \frac{\dd}{\dd t}h_\theta(z(t))
  =
  \sum_i\left|\frac{\partial}{\partial z^i}
  \bigl(\ee^{-i\theta}S\bigr)\right|^2
\]
in local holomorphic coordinates, while the Cauchy--Riemann equations give
\(\dd\varphi_\theta/\dd t=0\).  Thus each thimble is a constant-phase
steepest-descent cycle and the exponential decays at its ends by properness of
\(h_\theta\).  Standard finite-dimensional Morse theory for the pair
\(\{h_\theta\le H\}\subset X\), followed by \(H\to+\infty\), identifies the
middle-dimensional relative homology group with the free abelian group
generated by the upward thimbles.  The dual downward thimbles intersect them
with pairing \(\delta_{\sigma\tau}\), so the coefficient of
\(\mathcal J_\sigma^\theta\) in any relative cycle is
\(\langle\Gamma,\mathcal K_\sigma^\theta\rangle\).  This proves
\eqref{eq:thimble-homology-decomposition}, and integration of the same
holomorphic form over homologous relative cycles gives
\eqref{eq:thimble-integral-decomposition}.

For the expansion, apply the holomorphic Morse lemma near \(p_\sigma\):
\[
  S(z)=S(p_\sigma)+\frac12\sum_{i=1}^n \lambda_i w_i^2
\]
in local coordinates.  On the thimble the quadratic form has positive real
part after multiplication by \(\ee^{-i\theta}\).  Rescale
\(w_i=g^{1/2}y_i\), expand \(a(w)\) and the higher-coordinate corrections of
the volume form, and integrate term by term against the convergent Gaussian.
Odd monomials vanish in the diagonal Gaussian coordinates and even monomials
produce the powers in \eqref{eq:thimble-saddle-expansion}.  The square-root of
\(\det H_\sigma\) is the Gaussian determinant with the orientation and phase
chosen by the thimble.

A Stokes wall occurs precisely when two critical points can lie on a common
gradient trajectory, which requires equality of the constant phase
\(\varphi_\theta\).  Crossing such a wall changes the thimble basis by adding
integer multiples of thimbles connected by the flow.  Since this is a change
of basis in relative homology, the original cycle and integral are unchanged.
\end{proof}

\begin{remark}[Resurgence as a regulated statement]
\label{rem:resurgence-regulated-statement}
In finite dimensions the Borel singularities of the saddle expansion around
\(p_\sigma\) are controlled, under the usual analyticity and nondegeneracy
hypotheses, by action differences \(S(p_\tau)-S(p_\sigma)\).  Lateral Borel
sums across a Stokes ray differ by the contribution of thimbles whose critical
values become phase-aligned with \(p_\sigma\).  In a QFT path integral this
sentence becomes a theorem only after a regulator has turned the functional
integral into a finite-dimensional or otherwise rigorously controlled
analytic object, and after the regulator-removal limit has been proved in the
sector of interest.  Without those data, words such as ``instanton
resurgence'' or ``renormalon resurgence'' name a proposed analytic mechanism,
not a completed definition of the theory.
\end{remark}

\paragraph{Renormalons and Borel-plane obstructions.}
\label{rem:renormalons-borel-plane-obstructions}
The term \emph{renormalon} refers to a class of large-order effects in
renormalized perturbation theory whose diagnostic is a Borel singularity
associated with an integration over a momentum region.  It is not a theorem
that every perturbative series in a QFT has a renormalon of a specified type,
and it is not a replacement for a nonperturbative definition of the observable.
Nor should it be stated as a property of the beta function alone.  The
statement must always name the observable or coefficient function, the
expansion coordinate, the renormalization scheme, and the factorization
prescription.

\begin{proposition}[Factorial moment in a running-coupling model]
\label{prop:renormalon-factorial-moment-model}
For \(p>0\) and \(n\ge0\),
\[
  \int_0^1 x^{p-1}\bigl(-\log x\bigr)^n\,\dd x
  =
  \frac{n!}{p^{n+1}} .
  \label{eq:renormalon-factorial-moment}
\]
Consequently, if a controlled approximation to an observable contains
\[
  R(\alpha)
  =
  \alpha\int_0^1
  \frac{x^{p-1}\,\dd x}
       {1-\beta_0\alpha(-\log x)}
  \label{eq:renormalon-model-integral}
\]
and the geometric expansion is used before the integral is performed, then
\[
  R(\alpha)
  \sim
  \alpha\sum_{n\ge0}
  \frac{\beta_0^n n!}{p^{n+1}}\alpha^n,
  \qquad
  \mathcal B_R(u)=\frac{1}{p-\beta_0 u}
  \label{eq:renormalon-model-borel-pole}
\]
for the Borel transform of \(R(\alpha)/\alpha\).  The model therefore has a
positive-axis Borel pole at \(u=p/\beta_0\) when \(\beta_0>0\).
\end{proposition}

\begin{proof}
Set \(x=\ee^{-t}\).  Then \(x^{p-1}\dd x=\ee^{-pt}\dd t\) with \(t\in[0,\infty)\),
and
\[
  \int_0^1 x^{p-1}(-\log x)^n\,\dd x
  =
  \int_0^\infty \ee^{-pt}t^n\,\dd t
  =
  \frac{\Gamma(n+1)}{p^{n+1}}
  =
  \frac{n!}{p^{n+1}}.
\]
Expanding the denominator in
\eqref{eq:renormalon-model-integral} as
\(\sum_{n\ge0}[\beta_0\alpha(-\log x)]^n\) gives the displayed asymptotic
coefficients.  Dividing the coefficient of \(\alpha^{n+1}\) by \(n!\) gives
the geometric Borel series
\[
  \sum_{n\ge0}\frac{\beta_0^n}{p^{n+1}}u^n
  =
  \frac1{p-\beta_0 u}.
\]
\end{proof}

A controlled field-theoretic model calculation has the same algebraic core.
Suppose a Euclidean short-distance quantity contains, after a specified set
of approximations such as insertion of a chain of one-loop self-energy
bubbles, an integral over a momentum \(k\) of the form
\[
  \int \frac{\dd k}{k}\,F(k/Q)\,\alpha(k),
\]
where \(Q\) is the external scale, \(F\) restricts the relevant region, and
\(\alpha(k)\) is expanded around \(\alpha(Q)\) using one-loop running.  If
\[
  \alpha(k)
  =
  \frac{\alpha(Q)}
       {1+\beta_0\alpha(Q)\log(k/Q)}
  =
  \alpha(Q)\sum_{n\ge0}
  \bigl[-\beta_0\alpha(Q)\log(k/Q)\bigr]^n
\]
is inserted before the \(k\)-integral is performed, then integrals of powers of
\(\log(k/Q)\) can give \(n!\)-growth.  In the Borel transform this factorial
growth is represented, in this approximation, by singularities at finite Borel
parameter.

The name of the renormalon records the momentum region.  An infrared
renormalon is produced by the small-\(k\) region of such an integral; in an
asymptotically free theory such a singularity can lie on the positive Borel
ray for particular Euclidean short-distance quantities and obstruct an
unambiguous Borel sum of the perturbative coefficient function.  The associated
lateral-summation ambiguity then has the size of a power correction,
\[
  \exp[-c/\alpha(Q)]
  =
  \left(\frac{\Lambda_{\rm RG}}{Q}\right)^p
\]
for suitable constants \(c,p>0\), where \(\Lambda_{\rm RG}\) is the
renormalization-group invariant scale.  In QCD this is why infrared
renormalon analysis is tied to the operator product expansion: perturbation
theory for a Wilson coefficient and the definition of the corresponding
long-distance matrix element are not separately physical, and their
scheme/factorization ambiguities must cancel in the full observable.

\begin{proposition}[Algebraic form of OPE ambiguity cancellation]
\label{prop:ope-renormalon-ambiguity-cancellation}
Let \(Q\gg\mu\), and suppose a dimensionless Euclidean observable is represented
in a specified factorization scheme by the truncated operator product expansion
\[
  \mathcal X_\pm(Q)
  =
  C_{0,\pm}(Q,\mu)
  +
  \frac{1}{Q^p} C_p(Q,\mu) M_{p,\pm}(\mu)
  +
  O(Q^{-p-\delta})
  \label{eq:renormalon-ope-two-term-model}
\]
with \(p,\delta>0\).  Here \(C_{0,\pm}\) are the two lateral Borel sums of the
perturbative coefficient function for the identity operator, \(M_{p,\pm}\) are
the correspondingly defined matrix elements of an operator of mass dimension
\(p\), and \(C_p\) is assumed to have no lateral ambiguity at this order.  If
\[
  C_{0,+}(Q,\mu)-C_{0,-}(Q,\mu)
  =
  -
  \frac{1}{Q^p} C_p(Q,\mu)
  \bigl(M_{p,+}(\mu)-M_{p,-}(\mu)\bigr),
  \label{eq:ope-ambiguity-cancellation-condition}
\]
then
\[
  \mathcal X_+(Q)-\mathcal X_-(Q)=O(Q^{-p-\delta}).
  \label{eq:ope-ambiguity-cancellation-result}
\]
If the identity-coefficient ambiguity has the form
\[
  C_{0,+}-C_{0,-}
  =
  K(\mu)\left(\frac{\Lambda_{\rm RG}}{Q}\right)^p ,
  \label{eq:identity-coefficient-power-ambiguity}
\]
then cancellation at this order requires
\[
  M_{p,+}-M_{p,-}
  =
  -
  \frac{K(\mu)\Lambda_{\rm RG}^p}{C_p(Q,\mu)}
  \label{eq:operator-matrix-element-required-ambiguity}
\]
in the normalization used in \eqref{eq:renormalon-ope-two-term-model}.
\end{proposition}

\begin{proof}
Subtract the two lateral definitions of \(\mathcal X\) in
\eqref{eq:renormalon-ope-two-term-model}.  The remainder contributes only
\(O(Q^{-p-\delta})\), and the displayed leading difference is
\[
  C_{0,+}-C_{0,-}
  +
  \frac{1}{Q^p} C_p(M_{p,+}-M_{p,-}).
\]
This vanishes by \eqref{eq:ope-ambiguity-cancellation-condition}, proving
\eqref{eq:ope-ambiguity-cancellation-result}.  Substituting
\eqref{eq:identity-coefficient-power-ambiguity} into
\eqref{eq:ope-ambiguity-cancellation-condition} gives
\eqref{eq:operator-matrix-element-required-ambiguity}.  The proposition is
algebraic; a QFT application still has to construct the factorization scheme,
the operator \(M_p\), and the lateral sums.
\end{proof}

An ultraviolet renormalon is associated with the large-\(k\) region of the
corresponding expansion.  In theories whose perturbative coordinate grows
toward the ultraviolet, such as an IR-free scalar or Abelian gauge theory, this
large-momentum sensitivity is tied to the same short-distance loss of
perturbative control indicated by the Landau scale.  The precise location of a
renormalon singularity, and whether it lies on the integration ray relevant for
Borel summation, depends on the observable, the expansion variable, and the
renormalization scheme.  The invariant lesson is not that renormalons by
themselves define the theory, but that a perturbative representation can carry
ambiguities of order \(\exp[-c/\alpha]\), or \(\exp[-c/g^2]\) when \(g\) is the
gauge coupling.  Such terms are invisible to all orders in the perturbative
expansion and must be fixed, canceled, or interpreted by additional
nonperturbative or effective-field-theory data.
\footnote{For a systematic review, see M.~Beneke, ``Renormalons,''
\emph{Phys. Rept.} \textbf{317} (1999) 1--142,
doi:10.1016/S0370-1573(98)00130-6, arXiv:hep-ph/9807443.}

\paragraph{Application status ledgers.}
The preceding definitions are analytic objects, not decorative language.  A
QFT application of Borel summation, transseries, or Lefschetz thimbles must
name the regulated observable, the expansion coordinate, the analytic domain,
and the limiting statement being asserted.  The following datum is the
minimal bookkeeping used in this monograph.

\begin{definition}[Regulated asymptotic-analysis datum for a QFT observable]
\label{def:regulated-asymptotic-analysis-datum}
Let \(R\) label a regulator, let \(g\) be the perturbative expansion
coordinate, and let \(\mathcal X_R(g;\lambda)\) be a regulated observable,
where \(\lambda\) denotes all other named parameters, including masses,
subtraction scales, background sources, finite volume, and boundary
conditions.  A \emph{regulated asymptotic-analysis datum} consists of:
\begin{enumerate}
  \item a complex domain \(D_R\) in the \(g\)-plane on which
  \(\mathcal X_R(g;\lambda)\) is defined as a function or distributional
  matrix element;
  \item a formal perturbative series
  \[
    \widehat{\mathcal X}_R(g;\lambda)
    =
    \sum_{n\ge0}a_{R,n}(\lambda)g^n
  \]
  with the coefficient-extraction operation specified;
  \item a Borel transform or Borel--Leroy transform, together with its proved
  or hypothesized analytic continuation domain and growth bounds;
  \item, when nonperturbative sectors are present, a sector set \(S_R\), action
  differences \(A_{R,\sigma}(\lambda)\), powers \(\beta_{R,\sigma}\), and
  formal fluctuation series
  \[
    \ee^{-A_{R,\sigma}(\lambda)/g}g^{\beta_{R,\sigma}}
    \widehat\Phi_{R,\sigma}(g;\lambda);
  \]
  \item a statement of what happens as \(R\) is removed: coefficientwise limit,
  Borel-plane limit, distributional limit, correlation-function limit, or no
  continuum-limit claim.
\end{enumerate}
\end{definition}

\begin{remark}[Scope of a regulated asymptotic datum]
\label{rem:regulated-asymptotic-datum-status}
A datum in Definition~\ref{def:regulated-asymptotic-analysis-datum} proves a
Borel--Laplace reconstruction of \(\mathcal X_R\) in a direction \(\theta\)
only if the transform in item \(3\) satisfies the hypotheses of
Definition~\ref{def:gevrey-one-directional-borel-sum} and the resulting
Borel--Laplace function is identified with \(\mathcal X_R\) on a nonempty open
subdomain of \(D_R\).  It proves a QFT statement only after the asserted
regulator-removal limit in item \(5\) is established.  Without item \(5\), the
datum is a theorem or approximation for the regulated problem, not for the
continuum QFT.
\par
The first clause is exactly the content of the Borel--Laplace definition:
Gevrey growth supplies a locally convergent Borel transform, analytic
continuation and exponential bounds make the Laplace integral convergent, and
Theorem~\ref{thm:watson-borel-laplace} supplies the asymptotic expansion of
the resulting function.  The equality with the originally regulated
observable is an additional identity of analytic functions or distributions.
It follows from equality on a nonempty open set inside the common connected
domain by the identity theorem, or by equality of distributional boundary
values when the observable is defined distributionally.  Removing the regulator
is a separate limiting operation.  Thus a statement at fixed \(R\) cannot imply
a continuum-QFT statement unless item \(5\) is supplied.
\end{remark}

\begin{proposition}[Positive-ray Borel control in the stable quartic model]
\label{prop:zero-dimensional-quartic-positive-ray-borel-control}
For the zero-dimensional quartic integral of
Example~\ref{ex:zero-dimensional-quartic-large-order}, the Borel transform
\[
  B(\xi)
  =
  {}_2F_1\left(\frac14,\frac34;1;-\frac{2\xi}{3}\right)
\]
has no singularity for \(\xi\ge0\) and is bounded on the positive real
\(\xi\)-axis.  Therefore
\[
  \frac1g\int_0^\infty \ee^{-\xi/g}B(\xi)\,\dd\xi
\]
exists for every \(g>0\) and has the formal series
\(\sum_{n\ge0}a_ng^n\) of
\eqref{eq:zero-dimensional-quartic-coefficients} as its asymptotic expansion.
\end{proposition}

\begin{proof}
Use Euler's hypergeometric integral with
\[
  a=\frac14,\qquad b=\frac34,\qquad c=1,\qquad z=-\frac{2\xi}{3}.
\]
Since \(\operatorname{Re}c>\operatorname{Re}b>0\),
\[
  B(\xi)
  =
  \frac{1}{\Gamma(3/4)\Gamma(1/4)}
  \int_0^1
  t^{-1/4}(1-t)^{-3/4}
  \left(1+\frac{2\xi t}{3}\right)^{-1/4}\dd t .
  \label{eq:quartic-borel-euler-representation}
\]
For \(\xi\ge0\) the last factor is real, positive, and bounded above by
\(1\).  The remaining beta integral is finite.  Thus \(B\) is bounded on the
positive ray and has no positive-ray singularity.  The Borel--Laplace integral
therefore converges for \(g>0\), and
Theorem~\ref{thm:watson-borel-laplace} gives the displayed asymptotic
expansion because the Taylor coefficients of \(B\) are \(a_n/n!\), as computed
in Example~\ref{ex:zero-dimensional-quartic-large-order}.
\end{proof}

\begin{definition}[Quartic-oscillator spectral Borel datum]
\label{def:quartic-oscillator-spectral-borel-datum}
Let \(g\ge0\), and let \(H(g)\) be the Friedrichs realization on
\(L^2(\mathbb R,\dd x)\) of
\[
  H(g)
  =
  -\frac12\frac{\dd^2}{\dd x^2}
  +\frac12x^2
  +g x^4 .
  \label{eq:quartic-oscillator-hamiltonian}
\]
The domain is determined by the closed quadratic form
\[
  q_g[\psi]
  =
  \frac12\|\psi'\|_{L^2}^2
  +\frac12\|x\psi\|_{L^2}^2
  +g\|x^2\psi\|_{L^2}^2 .
\]
Its spectrum is discrete and simple; write \(E_n(g)\), \(n=0,1,\ldots\), for
the increasing eigenvalues.  The Rayleigh--Schr\"odinger perturbative series at
the \(n\)-th harmonic-oscillator level is
\[
  \widehat E_n(g)=\sum_{r\ge0}E_{n,r}g^r .
  \label{eq:quartic-oscillator-perturbation-series}
\]
A Borel-summability statement for
\eqref{eq:quartic-oscillator-perturbation-series} means that this particular
spectral branch \(E_n(g)\) is recovered by a Borel--Laplace transform
satisfying
Definition~\ref{def:gevrey-one-directional-borel-sum}.
\end{definition}

\begin{quotedtheorem}[Quartic-oscillator Borel summability]
\label{qthm:quartic-oscillator-borel-summability}
For each fixed \(n\), the series
\eqref{eq:quartic-oscillator-perturbation-series} is Borel summable on the
positive real \(g\)-axis to the eigenvalue \(E_n(g)\) of
\eqref{eq:quartic-oscillator-hamiltonian}.  More precisely, the perturbative
coefficients are Gevrey-one, the Borel transform has analytic continuation
and growth in a sector containing the positive Borel ray, and the
Borel--Laplace integral in that direction equals \(E_n(g)\) for \(g>0\).
\end{quotedtheorem}

\begin{proof}[Status of the quoted theorem]
This is a theorem of Graffi--Grecchi--Simon for even polynomial anharmonic
oscillators.  Its proof uses sectorial analyticity and resolvent estimates
for the operator family \(H(g)\), not diagram counting.  The theorem is
quoted here as a well-established one-dimensional spectral result rather than
reproved, because its proof belongs to analytic perturbation theory of
ordinary differential operators.  The monograph uses it only as a controlled
comparison example: in QFT the corresponding estimates must be supplied for
the regulated field-theoretic observable and then passed through the
regulator-removal limit.
\end{proof}

\begin{proposition}[Local Borel convergence from Gevrey-one growth]
\label{prop:coefficient-growth-not-borel-summability}
Suppose a formal series \(\sum_{r\ge0}a_rg^r\) obeys
\(|a_r|\le C M^r r!\).  This bound alone implies convergence of the formal
Borel transform near \(\xi=0\), but it does not imply Borel summability of a
named observable.
\end{proposition}

\begin{proof}
The estimate gives
\[
  \left|\frac{a_r}{r!}\xi^r\right|
  \le C(M|\xi|)^r ,
\]
so the Borel series converges for \(|\xi|<M^{-1}\).  Definition
\ref{def:gevrey-one-directional-borel-sum} requires more: analytic
continuation of the Borel transform to a sector containing the summation ray,
an exponential-growth bound there, existence of the Laplace integral, and an
identity between that Laplace integral and the observable on a nonempty
domain.  None of these follows from the local coefficient bound.  Thus
large-order estimates are evidence for a possible summation mechanism, but
not the mechanism itself.
\end{proof}

\begin{definition}[Instanton-sector transseries ledger]
\label{def:instanton-sector-transseries-ledger}
Consider a regulated Euclidean gauge-theory path integral whose configuration
space has connected components indexed by an integer topological charge
\(k\in\mathbb Z\).  Assume the Euclidean action in sector \(k\) has the form
\[
  S_{E,R}[A,\Psi]
  =
  \frac{1}{4g^2}\int \operatorname{tr}_\Delta(F\wedge *F)
  -\ii\theta k
  +S_{R,\mathrm{matter}}[A,\Psi],
  \label{eq:instanton-ledger-action}
\]
where \(\operatorname{tr}_\Delta\) is the trace normalization fixed in the
gauge-theory convention dictionary of this monograph.  An instanton-sector
transseries ledger for a regulated observable is a formal expansion
\[
  \widehat{\mathcal X}_R(g,\theta)
  =
  \sum_{k\in\mathbb Z}
  \sum_{\alpha\in\mathcal S_{R,k}}
  \ee^{-A_{R,k,\alpha}/g^2}\,
  \ee^{\ii k\theta}\,
  g^{\beta_{R,k,\alpha}}\,
  \widehat\Phi_{R,k,\alpha}(g),
  \label{eq:instanton-sector-transseries-ledger}
\]
where \(\alpha\) labels the regulated saddle or quasi-saddle data inside the
charge-\(k\) sector.  The notation includes the fluctuation determinant,
collective-coordinate measure, zero-mode insertions, and boundary
contributions only if these have been specified by the regulator.
\end{definition}

\begin{proposition}[Algebraic constraints on an instanton-sector ledger]
\label{prop:instanton-ledger-algebra}
For a ledger of the form
\eqref{eq:instanton-sector-transseries-ledger}, the \(\theta\)-dependence is
\(2\pi\)-periodic.  If the regulated observable is real at \(\theta=0\) and
the sector data obey the conjugation rule
\[
  A_{R,-k,\bar\alpha}=A_{R,k,\alpha},
  \qquad
  \beta_{R,-k,\bar\alpha}=\beta_{R,k,\alpha},
  \qquad
  \widehat\Phi_{R,-k,\bar\alpha}(g)
  =
  \overline{\widehat\Phi_{R,k,\alpha}(g)}
  \]
for \(g>0\), then the paired \(k\) and \(-k\) contribution at real \(\theta\)
is real and equals
\[
  2\,\operatorname{Re}
  \left[
    \ee^{-A_{R,k,\alpha}/g^2}
    \ee^{\ii k\theta}
    g^{\beta_{R,k,\alpha}}
    \widehat\Phi_{R,k,\alpha}(g)
  \right].
  \label{eq:instanton-pair-real-combination}
\]
\end{proposition}

\begin{proof}
Since \(k\in\mathbb Z\),
\[
  \ee^{\ii k(\theta+2\pi)}=\ee^{\ii k\theta}.
\]
This proves the periodicity.  The stated conjugation rule makes the
\(-k\)-sector term the complex conjugate of the \(k\)-sector term for real
\(\theta\) and \(g>0\).  Adding the two terms gives twice the real part,
which is \eqref{eq:instanton-pair-real-combination}.
\end{proof}

\begin{proposition}[BPST sector in the asymptotic-analysis ledger]
\label{prop:bpst-sector-asymptotic-ledger}
In the trace-delta Yang--Mills convention of this monograph,
\[
  S_{\rm YM,E}
  =
  \frac{1}{4g_{\rm YM}^2}
  \int_{\mathbb R^4}\operatorname{tr}_\Delta
  (F_{\mu\nu}F_{\mu\nu})\,\dd^4x
  -\ii\theta Q[A],
  \label{eq:bpst-ledger-action-trace-delta}
\]
the \(SU(2)\) BPST one-instanton sector contributes the exponential ledger
factor
\[
  \exp\left(-\frac{4\pi^2}{g_{\rm YM}^2}+\ii\theta\right)
  \label{eq:bpst-ledger-factor-trace-delta}
\]
to any regulated observable whose insertions have nonzero projection onto
that sector.  In the common half-trace convention
\(\operatorname{tr}(T_aT_b)=\delta_{ab}/2\), this same factor is written
\(\exp[-8\pi^2/g_{\rm ht}^2+\ii\theta]\), with
\[
  g_{\rm ht}=\sqrt2\,g_{\rm YM}.
  \label{eq:half-trace-trace-delta-coupling-relation-ledger}
\]
\end{proposition}

\begin{proof}
The explicit BPST calculation in
Proposition~\ref{prop:bpst-one-instanton-data} gives \(Q[A]=1\) and, in the
active trace-delta convention,
\[
  \frac{1}{4g_{\rm YM}^2}
  \int_{\mathbb R^4}
  \operatorname{tr}_\Delta(F_{\mu\nu}F_{\mu\nu})\,\dd^4x
  =
  \frac{4\pi^2}{g_{\rm YM}^2}.
\]
Substitution in \eqref{eq:bpst-ledger-action-trace-delta} gives
\eqref{eq:bpst-ledger-factor-trace-delta}.  The half-trace basis used in much
of the instanton literature differs from the trace-delta basis by
\(t_a=\sqrt2\,T_a\).  Therefore the same connection is represented with
\(g_{\rm ht}^2=2g_{\rm YM}^2\), and
\[
  \frac{8\pi^2}{g_{\rm ht}^2}
  =
  \frac{4\pi^2}{g_{\rm YM}^2}.
\]
The equality is only a convention change; it does not alter the topological
charge or the physical semiclassical weight.
\end{proof}

\begin{controlledapproximation}[Semiclassical instanton expansion]
\label{capprox:semiclassical-instanton-expansion-status}
The factor \eqref{eq:bpst-ledger-factor-trace-delta} is not yet an instanton
calculus formula for a continuum QFT observable.  A semiclassical
one-instanton approximation additionally requires a regulator, a gauge slice
or BV treatment of the zero modes, the determinant ratio, the collective
coordinate density on the instanton moduli space, insertion zero-mode
saturation, and a prescription for boundary strata such as small instantons.
Only after those data are defined and their regulator-removal behavior is
proved does the ledger become a theorem about the observable rather than a
semiclassical sector label.
\end{controlledapproximation}

\begin{remark}[Status of instanton and renormalon applications]
\label{rem:instanton-renormalon-application-status}
Definition~\ref{def:instanton-sector-transseries-ledger} is a bookkeeping
standard, not an assertion that the instanton expansion has been justified.
For a gauge theory it remains necessary to define the path integral regulator,
the topological sectors, the treatment of gauge zero modes and collective
coordinates, the determinant normalization, the behavior near small-size
instantons or other boundary strata, and the continuum limit.  Renormalon
ambiguities are different objects: they arise from momentum-region
singularities in perturbative coefficient functions and are tied to
factorization and operator definitions.  Instanton saddles and renormalon
singularities can coexist in the same observable, but neither one is a
substitute for the other.
\end{remark}

In this monograph a displayed perturbative beta function or critical exponent
has the following default meaning unless a stronger statement is stated: it is
a coefficientwise construction in a specified regularized and renormalized
chart, giving an asymptotic small-coupling expansion or an effective
finite-order approximation in its stated domain.  A finite truncation can be
useful and sharply predictive when the running coupling and logarithms are
controlled.  A nonperturbative definition of the QFT, or a convergence theorem
for the infinite formal series, requires additional data beyond the finite
truncation.

\section{Invariant Data at Fixed Points}

The preceding one-coordinate computation is a special case of the tensorial
transformation law for a vector field.  Let \(\mathcal T\) be a finite
dimensional coordinate patch in the renormalized theory family, with local
coordinates \(\lambda^I\).  The beta function is the vector field
\[
  \beta=\beta^I(\lambda)\frac{\partial}{\partial\lambda^I}.
\]
A finite scheme change is a local diffeomorphism
\[
  \widetilde\lambda^A=f^A(\lambda),
  \qquad
  J^A{}_I(\lambda)=\frac{\partial f^A}{\partial\lambda^I}.
\]
The components of the same vector field in the new coordinates are
\[
  \widetilde\beta^A(\widetilde\lambda)
  =
  J^A{}_I(\lambda)\,\beta^I(\lambda),
  \qquad
  \widetilde\lambda=f(\lambda).
\]
Thus a beta function is not a coordinate-independent list of functions; the
coordinate-independent object is the flow it defines on \(\mathcal T\).

A fixed point is a zero of this vector field:
\[
  \beta^I(\lambda_\ast)=0 .
\]
This statement is invariant under changes of coordinates.  Let
\[
  M^I{}_J
  =
  \left.
  \frac{\partial\beta^I}{\partial\lambda^J}
  \right|_{\lambda_\ast}
\]
be the linearization of the RG vector field at the fixed point.  Since the
extra term obtained by differentiating \(J^A{}_I(\lambda)\beta^I(\lambda)\)
is proportional to \(\beta(\lambda_\ast)\), it vanishes at the fixed point.
Therefore
\[
  \widetilde M^A{}_B
  =
  J^A{}_I(\lambda_\ast)\,
  M^I{}_J\,
  (J^{-1})^J{}_B(\lambda_\ast).
\]
The linearized matrices in two schemes are similar.  Their eigenvalues are
therefore invariant, although their eigenvectors are represented by different
coordinate components.  These eigenvalues are the critical exponents of the
linearized flow in the chosen normalization of RG time.

If \(M\) has zero eigenvalues, the linearized analysis identifies tangent
directions that require higher-order study.  A zero eigenvalue is the
condition for linearized marginality; exact marginality is the stronger
statement that the RG vector field vanishes along an actual submanifold of
theory space.  In local QFT this statement must also be made after quotienting
redundant directions generated by field redefinitions and after imposing the
symmetries used to define the coordinate patch.

Figure~\ref{fig:volume-ii-scheme-change} records the coordinate-geometric
content of a finite scheme change: the same renormalized theory family is
described by different finite coordinates, and the beta function transforms as
a vector field.

\begin{figure}[H]
  \centering
  \resizebox{0.76\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[draw=black, rounded corners, align=center, inner sep=7pt]
      (A) at (-2.8,0) {coordinate \(\lambda\)\\\(\beta(\lambda)\)};
    \node[draw=black, rounded corners, align=center, inner sep=7pt]
      (B) at (2.8,0) {coordinate \(\widetilde\lambda\)\\
      \(\widetilde\beta(\widetilde\lambda)\)};
    \draw[-{Latex[length=2mm]}, thick] (A) -- node[above]
      {\(\widetilde\lambda=f(\lambda)\)} (B);
    \draw[-{Latex[length=2mm]}, thick] (B) to[bend left=18]
      node[below] {\(\beta=\widetilde\beta\circ f/f'\)} (A);
    \node[align=center, blue!70!black] at (0,-1.55)
      {same renormalized family\\different finite coordinates};
  \end{tikzpicture}
  }
  \caption{A finite scheme change is a change of coordinates on the
  renormalized theory family.  The beta function transforms as a vector field
  under this change of coordinate.}
  \label{fig:volume-ii-scheme-change}
\end{figure}

The next chapter applies the same scale-dependent logic to local operator
insertions.  Infinitesimal deformations of the action define renormalized
operators, and their scale dependence is governed by the linearization of the
beta-function vector field.
