\chapter{Global Anomalies, Spontaneous Symmetry Breaking, and Pions}
\label{ch:volume-ii-21}
\ConventionsLorentzian


The four-dimensional orientation convention is the one used in the local
anomaly calculation of Chapter~\ref{ch:volume-ii-20}:
\[
  \epsilon^{0123}=+1,
  \qquad
  \widetilde F^{\mu\nu}
  =
  \frac12\epsilon^{\mu\nu\rho\sigma}F_{\rho\sigma}
\]
for any two-form \(F\).  This convention is used for the perturbative triangle
contact terms, Wess--Zumino--Witten functional, and electromagnetic
specialization below.

\section{Background Fields and Anomaly Data}

Let a quantum field theory have an exact internal global symmetry
\(G_{\mathrm f}\).  The symmetry acts on local fields and local operators, and
on Hilbert spaces obtained after choosing a phase.  To test the symmetry as a
property of local correlation functions, introduce a nondynamical
\(G_{\mathrm f}\)-connection \(B\) on spacetime \(M\).  The background-field
generating functional is
\[
  Z[B]
  =
  \int [D\Phi]\,
  \exp\{\ii S[\Phi;B]\},
\]
where \(\Phi\) denotes the dynamical fields and \(B\) is held fixed as an
external datum.
Let \(\mathcal B\) be the space of allowed background fields and
\(\mathcal G_{\mathrm f}\) the group of background gauge transformations.  If
\(g\in\mathcal G_{\mathrm f}\), write \(B^g\) for the transformed connection.

\begin{definition}[Background anomaly cocycle]
\label{def:background-anomaly-cocycle}
The background anomaly is the equivalence class, under local counterterms, of
phases
\(\mathcal A_g[B]\) in the transformation law
\begin{equation}
  Z[B^g]
  =
  \exp\{\ii\mathcal A_g[B]\}\,Z[B],
  \label{eq:finite-background-anomaly}
\end{equation}
where \(g\in\mathcal G_{\mathrm f}\).  The zero class is the class for which
there exists a local counterterm \(C[B]\) making the transformed functional
invariant under all background gauge transformations.  Multiplying \(Z[B]\) by
\(\exp\{\ii C[B]\}\) changes the representative by
\[
  \mathcal A_g[B]
  \longmapsto
  \mathcal A_g[B]+C[B^g]-C[B].
\]
Thus the anomaly is represented by the equivalence class of
\(\mathcal A_g[B]\) under local-counterterm shifts.
\end{definition}

\begin{remark}[Finite cocycle and Wess--Zumino consistency]
\label{rem:wess-zumino-consistency-from-cocycle}
The phases in \eqref{eq:finite-background-anomaly} obey a cocycle identity.
Applying \(h\) and then \(g\) gives the same transformed background as applying
\(gh\), hence
\begin{equation}
  \mathcal A_{gh}[B]
  =
  \mathcal A_g[B^h]+\mathcal A_h[B]
  \quad
  \operatorname{mod} 2\pi .
  \label{eq:background-anomaly-cocycle}
\end{equation}
The infinitesimal part of this cocycle is the local Wess--Zumino consistency
condition developed in the preceding chapter.  A finite anomaly may also have a
purely finite part: its differential on the identity component of
\(\mathcal G_{\mathrm f}\) vanishes, but the cocycle
\eqref{eq:background-anomaly-cocycle} is nontrivial on another component or on
a noncontractible loop in background-field space.

Equation \eqref{eq:finite-background-anomaly} gives
\[
  Z[B^{gh}]
  =
  \exp\{\ii\mathcal A_{gh}[B]\}Z[B].
\]
The same background is obtained by first applying \(h\) and then applying
\(g\):
\[
  Z[(B^h)^g]
  =
  \exp\{\ii\mathcal A_g[B^h]\}Z[B^h]
  =
  \exp\{\ii\mathcal A_g[B^h]+\ii\mathcal A_h[B]\}Z[B].
\]
The two answers differ by an integer multiple of \(2\pi\) in the exponent,
which is \eqref{eq:background-anomaly-cocycle}.  Differentiating the finite
identity at the identity element of \(\mathcal G_{\mathrm f}\) gives the
infinitesimal consistency condition for the corresponding local anomaly
representative.
\end{remark}

Geometrically, the background-field construction assigns a complex line to each
gauge-equivalence class of \(B\), and \(Z[B]\) is a section of this anomaly
line.  Local counterterms change a trivialization of the line.  When the anomaly is
represented by an invertible field theory on a five-dimensional manifold \(Y\)
with \(\partial Y=M\), equality of boundary variations identifies the phase in
\eqref{eq:finite-background-anomaly}.  The local Chern--Simons version of this
bulk-boundary identity was described in Chapter~\ref{ch:volume-ii-20}; below
we use mapping tori to state the finite version needed for global anomalies.

Figure~\ref{fig:anomaly-matching-backgrounds} records the anomaly-matching
statement in background-field language: microscopic and infrared descriptions
are compared as sections of the same anomaly line over background gauge fields.

\begin{figure}[H]
  \centering
  \resizebox{0.92\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[align=center] (uv) at (-4.4,1.35)
      {microscopic theory\\with background \(B\)};
    \node[align=center] (ir) at (4.4,1.35)
      {infrared description\\with the same \(B\)};
    \node[align=center] (uvg) at (-4.4,-1.35)
      {transformed background\\\(B^g\)};
    \node[align=center] (irg) at (4.4,-1.35)
      {same anomaly phase\\in the infrared};
    \draw[-{Latex[length=2mm]}, thick] (uv)--node[above]
      {RG flow} (ir);
    \draw[-{Latex[length=2mm]}, thick] (uvg)--node[below]
      {RG flow} (irg);
    \draw[-{Latex[length=2mm]}, blue!70!black, thick] (uv)--node[left]
      {\(g\)} (uvg);
    \draw[-{Latex[length=2mm]}, blue!70!black, thick] (ir)--node[right]
      {\(g\)} (irg);
    \node[align=center, blue!70!black] at (0,-0.05)
      {\(\exp\{\ii\mathcal A_g[B]\}\) is unchanged\\modulo local counterterms};
  \end{tikzpicture}}
  \caption{Anomaly matching is equality of background-response classes.  The
  ultraviolet and infrared functionals may use different fields, but their
  transformation cocycles agree after local counterterm conventions are
  identified.}
  \label{fig:anomaly-matching-backgrounds}
\end{figure}

\section{Anomaly Matching}

\begin{remark}[Anomaly matching as equality of background-response classes]
\label{rem:anomaly-matching-background-cocycle-class}
Let a microscopic theory with exact global symmetry \(G_{\mathrm f}\) flow to
an infrared description.  Couple both descriptions to the same background
\(B\).  The endpoint statement called anomaly matching is that, if the
ultraviolet generating functional transforms as in
\eqref{eq:finite-background-anomaly}, then the infrared functional obeys
\[
  Z_{\mathrm{IR}}[B^g]
  =
  \exp\{\ii\mathcal A_g[B]\}\,Z_{\mathrm{IR}}[B]
\]
after possible local counterterms have been matched.  The equality of the
equivalence class of \(\mathcal A_g[B]\) modulo counterterms is 't Hooft
anomaly matching.

The matching object is the cocycle class
\([\mathcal A]\) defined by \eqref{eq:background-anomaly-cocycle}.  Different
infrared descriptions may carry the same class: massless fermions can produce
it through their determinant or Pfaffian line, Goldstone fields can produce it
through a Wess--Zumino--Witten functional, and a topological sector can produce
it through an invertible response.  The noncircular statement is not that the
full ultraviolet and infrared partition functions differ by a background-only
local factor.  It is that, at finite regulator, Wilsonian integration is a
pushforward of a section of the same anomaly line over the same background
groupoid.

Fix a regulator \(\Lambda\), a lower scale \(\Lambda'\), and a
background-equivariant split of fields
\(\Phi=(\varphi,\chi)\), where \(\chi\) denotes the shell to be removed.  In a
chosen local frame for the anomaly line, define the Wilsonian shell
pushforward by
\begin{equation}
  \exp\{\ii S_{\Lambda'}[\varphi;B]\}
  =
  \int_{\chi\in\mathrm{shell}(\Lambda,\Lambda';B,\varphi)}
  [D\chi]_{\Lambda,B,\varphi}\,
  \exp\{\ii S_{\Lambda}[\varphi,\chi;B]\}.
  \label{eq:finite-regulator-anomaly-shell-pushforward}
\end{equation}
The shell domain and measure are part of the regulator datum.  Assume they are
mapped to the corresponding shell for \(B^g\) under the microscopic
background-gauge transformation.  If the finite-regulator microscopic measure
and action transform by
\[
  [D\Phi]_{\Lambda,B^g}\exp\{\ii S_\Lambda[\Phi^g;B^g]\}
  =
  \exp\{\ii\mathcal A^\Lambda_g[B]\}\,
  [D\Phi]_{\Lambda,B}\exp\{\ii S_\Lambda[\Phi;B]\},
\]
then changing variables \(\chi^g\mapsto\chi\) in
\eqref{eq:finite-regulator-anomaly-shell-pushforward} gives
\begin{equation}
  \exp\{\ii S_{\Lambda'}[\varphi^g;B^g]\}
  =
  \exp\{\ii\mathcal A^\Lambda_g[B]\}\,
  \exp\{\ii S_{\Lambda'}[\varphi;B]\}
  \label{eq:finite-regulator-shell-inherits-anomaly}
\end{equation}
in the same line frame, up to the variation of explicitly inserted local
counterterms.  If the retained-field measure frame is changed, the additional
Jacobian is recorded as a local retained-field coboundary
\(K[\varphi^g;B^g]-K[\varphi;B]\).  Thus the retained-field Wilsonian
functional is generally not a background-only factor; it is a new section whose
line transport cocycle is the old one.  This is the finite-cutoff content of
anomaly preservation.

The locality needed for Wilsonian physics is a separate hypothesis.  When the
removed shell is separated from the retained momenta and no light threshold is
integrated out, the difference between two nearby shell prescriptions is
quasi-local in \(\varphi\) and \(B\).  Its background variation is a local
counterterm coboundary in the anomaly-line frame.  Quasi-locality is enough for
the anomaly class because local counterterms change
\(\mathcal A_g[B]\) only by
\[
  C[B^g]-C[B]
\]
or by the corresponding local retained-field counterterm before the remaining
fields are integrated.  It is not a claim that the full one-particle-irreducible
functional, or the final infrared partition function after massless fields have
been integrated, is local.

Passing from this finite-regulator statement to endpoint anomaly matching is a
conditional structural principle.  One must assume that the microscopic theory
and the proposed infrared description are obtained as compatible continuum and
long-distance limits of sections over the same background groupoid, with the
same exact symmetry rather than an accidentally enlarged infrared one.  One
must also specify how massless sectors, Goldstone fields, topological order, or
invertible responses provide the infrared section or boundary condition for
the same anomaly line.  Under these assumptions the anomaly-line class
transported by finite shell pushforwards is the class compared between the
ultraviolet theory and the infrared description.  For local perturbative
anomalies this is the descent class; for flat global anomalies it is the
determinant/Pfaffian-line holonomy around loops in the background groupoid, as
in Chapters~\ref{ch:global-anomaly-inflow-invertible-field-theories} and
\ref{ch:curved-eta-invariants-global-anomalies}.
\end{remark}

\section{A Perturbative Example of a Global Anomaly}

The local triangle computation already distinguishes two uses of a current.
Consider a four-dimensional massless Dirac fermion with vector and axial
currents
\[
  j^\mu=\ii\bar\psi\gamma^\mu\psi,
  \qquad
  j_A^\mu=\ii\bar\psi\gamma^\mu\gamma_5\psi .
\]
The preceding chapter fixed the regularization convention in which the vector
current coupled to an ordinary \(U(1)\) gauge field is preserved and the axial
current carries the \(j_Ajj\) anomaly.  To test the axial \(U(1)\) itself as a
background symmetry, couple a background gauge field \(A_\mu\) to \(j_A^\mu\),
\[
  \Delta S=\int\dd^4x\,A_\mu j_A^\mu .
\]
Gauge invariance of this background coupling would require the Ward identity
for the vector-current divergence in the three-point function with two axial
insertions:
\[
  \partial_\mu
  \bra{\vac}
  \operatorname T\,j^\mu(x)j_A^\nu(y)j_A^\rho(z)
  \ket{\vac} .
\]
The same oriented triangle graphs as in the local axial anomaly appear, with
two \(\gamma_5\) insertions.  In the \(D=4-\epsilon\) regulator, split
\(\ell=\ell_\parallel+\ell_\perp\) and keep \(\gamma_5\) four-dimensional.
The trace term that remains after imposing the two axial-vector external Ward
identities is proportional to
\[
  \ell_\perp^2(-4\ii)\epsilon^{\nu\rho\alpha\beta}
  (k-2\ell)_\alpha p_\beta .
\]
The loop integral is symmetric under permutations of the three external
momenta \(k,p,-k-p\).  Its finite local part is linear in the external
momenta, hence
\[
  \epsilon^{\nu\rho\alpha\beta}(k-2\ell)_\alpha p_\beta
  \longmapsto
  \frac13\epsilon^{\nu\rho\alpha\beta}k_\alpha p_\beta
\]
inside the regulated contact term.  With the conventions of the preceding
chapter this gives
\begin{equation}
  \partial_\mu
  \bra{\vac}
  \operatorname T\,j^\mu(x)j_A^\nu(y)j_A^\rho(z)
  \ket{\vac}
  =
  -\frac1{6\pi^2}
  \epsilon^{\nu\rho\alpha\beta}
  \partial_\alpha\delta^{(4)}(x-z)
  \partial_\beta\delta^{(4)}(y-z).
  \label{eq:vector-current-two-axial-contact-term}
\end{equation}
Expanding the background-field effective action
\[
  \ee^{\ii\Gamma[A]}
  =
  \int D\psi\,
  \exp\left\{\ii S_0[\psi]
    +\ii\int\dd^4x\,A_\mu j_A^\mu\right\}
\]
to second order in \(A\), equation
\eqref{eq:vector-current-two-axial-contact-term} implies
\begin{equation}
  \Gamma[A+\dd\zeta]-\Gamma[A]
  =
  \int\dd^4x\,\zeta(x)\,
  \frac1{48\pi^2}
  \epsilon^{\mu\nu\rho\sigma}
  F_{\mu\nu}F_{\rho\sigma}.
  \label{eq:axial-background-gauge-variation}
\end{equation}

\begin{proposition}[Axial background obstruction]
\label{prop:axial-background-obstruction}
In the vector-current-preserving regularization convention used above, the
massless Dirac fermion has a nontrivial anomaly for the axial \(U(1)\)
background.  More precisely, no four-dimensional local counterterm depending
only on the axial background \(A\), within this convention, can make
\(\Gamma[A]\) invariant under all compactly supported infinitesimal
background transformations \(A\mapsto A+\dd\zeta\).  Therefore the axial
background cannot be promoted to an ordinary dynamical gauge field unless
additional matter or topological data cancels the class represented by
\eqref{eq:axial-background-gauge-variation}.
\end{proposition}

\begin{proof}
Gauge invariance of an Abelian background functional would require
\(\delta_\zeta\Gamma[A]=0\) for every test function \(\zeta\).  Equation
\eqref{eq:axial-background-gauge-variation} gives instead
\[
  \delta_\zeta\Gamma[A]
  =
  \frac1{48\pi^2}
  \int_{M_4}\zeta\,\epsilon^{\mu\nu\rho\sigma}
  F_{\mu\nu}F_{\rho\sigma}.
\]
Since \(F\wedge F\) can be varied independently in a compact spacetime
region, this distribution is not identically zero.  A local counterterm would
change the right hand side by a local Abelian Wess--Zumino coboundary.  In the
single-background Abelian complex there is no four-form local functional
whose infinitesimal variation is \(\zeta\,F\wedge F\): the only possible
Chern--Simons primitive \(A\wedge F\) is a three-form, while \(A\wedge
A\wedge F\) vanishes for an Abelian one-form.  Equivalently, the class
whose descent gives \(\zeta F\wedge F\) is the nonzero six-form anomaly
polynomial proportional to \(F\wedge F\wedge F\).  Thus the obstruction
cannot be removed without changing the theory, changing the symmetry being
gauged, or adding another sector with the opposite anomaly.
\end{proof}

Thus the axial rotation is a symmetry of the classical massless action in
flat space without background flux, but its quantum coupling to an axial
background is anomalous: the partition function is naturally a section of an
anomaly line rather than a function on the quotient by axial background gauge
transformations.

\section{Example: A Finite SU(2) Anomaly}

Let \((M,\sigma_M)\) be a compact spin four-manifold and let
\(G_{\mathrm f}=SU(2)\).  The spin structure \(\sigma_M\) is part of the
background datum: a Weyl fermion is not defined on a merely oriented
four-manifold.  A change of \(\sigma_M\), when several spin structures exist,
is a change of the fermionic QFT background and may change the Pfaffian line
and its anomaly.  Consider one left-handed Weyl fermion valued in the
fundamental representation of \(SU(2)\).  For an \(SU(2)\) background
connection \(B\), the Euclidean Weyl operator is
\[
  D_{B,\sigma_M}^+:
  \Gamma(S_+\otimes E_B)
  \longrightarrow
  \Gamma(S_-\otimes E_B),
\]
where \(S_\pm\) are the positive and negative chirality spin bundles and
\(E_B\) is the associated rank-two complex vector bundle.  The fundamental
representation of \(SU(2)\) is pseudoreal, so the fermion functional integral
defines a Pfaffian line over the space of structured backgrounds
\((M,\sigma_M,B)\).

Let \(g:M\to SU(2)\) be a background gauge transformation, and choose a smooth
path of background connections \(B_t\), \(t\in[0,1]\), from \(B_0=B\) to
\(B_1=B^g\).  Identifying the two ends of \(M\times[0,1]\) by \(g\) gives the
mapping torus
\[
  X_g
  =
  (M\times[0,1])/\bigl((x,1)\sim(x,0)\hbox{ by }g\bigr)
\]
with an induced \(SU(2)\) bundle \(E_g\).  The spin structure on
\(M\times[0,1]\) is the product one from \(\sigma_M\); gluing the two ends
by the identity diffeomorphism of \(M\) and the internal gauge transformation
\(g\) gives a spin structure \(\sigma_g\) on \(X_g\).  Thus the invariant
below is a function of \((X_g,\sigma_g,E_g)\), not just of the oriented
manifold and the vector bundle.  If the spin structure on \(M\) is changed, or
if the mapping-torus spin structure is twisted by an allowed
\(\mathbb Z_2\) flat line, the mod-two index may change.

Let \(\mathcal D_{X_g,\sigma_g,E_g}\) be the five-dimensional real Dirac
operator coupled to this bundle, and define its mod-two index by
\[
  \zeta(X_g,\sigma_g,E_g)
  =
  \dim\ker\mathcal D_{X_g,\sigma_g,E_g}
  \quad \operatorname{mod}2 .
\]
The analytic framework behind this statement is the Dai--Freed formulation of
fermion anomalies.  For a family of chiral Dirac operators, the determinant or
Pfaffian is a section of a line over background-field space, and the holonomy
of this line around a closed loop of backgrounds is computed by the
Atiyah--Patodi--Singer eta-invariant of the Dirac operator on the associated
mapping torus.  In the present pseudoreal \(SU(2)\) fundamental case the real
structure refines the Dai--Freed holonomy to the mod-two-index sign
\begin{equation}
  \operatorname{Hol}_{\mathrm{DF}}(X_g,\sigma_g,E_g)
  =
  (-1)^{\zeta(X_g,\sigma_g,E_g)} .
  \label{eq:su2-dai-freed-holonomy}
\end{equation}
Equivalently, cutting the mapping torus open gives the APS-boundary
description of the same Pfaffian transport.  Therefore the Pfaffian transforms
by
\begin{equation}
  \frac{\operatorname{Pf}(D_{B^g,\sigma_M}^+)}
       {\operatorname{Pf}(D_{B,\sigma_M}^+)}
  =
  (-1)^{\zeta(X_g,\sigma_g,E_g)} .
  \label{eq:su2-global-anomaly-pfaffian}
\end{equation}
For \(M=S^4\) and \(g\) representing the nontrivial class in
\(\pi_4(SU(2))\simeq\mathbb Z_2\), the five-dimensional mod-two index theorem
gives \(\zeta(X_g,\sigma_g,E_g)=1\) for one fundamental Weyl doublet.  The
following proposition is the QFT consequence of this index-theoretic input.

\begin{proposition}[Finite \(SU(2)\) anomaly criterion for doublets]
\label{prop:finite-su2-anomaly-doublets}
Let \(n\) left-handed Weyl fermions transform as fundamental doublets of a
background \(SU(2)\) symmetry on a compact spin four-manifold
\((M,\sigma_M)\).  For a closed loop of backgrounds represented by the mapping
torus \((X_g,\sigma_g,E_g)\), the product Pfaffian line has holonomy
\[
  \operatorname{Hol}(X_g,\sigma_g,E_g)
  =
  (-1)^{n\,\zeta(X_g,\sigma_g,E_g)} .
\]
For the test family \(M=S^4\) with \(g\) the generator of
\(\pi_4(SU(2))\simeq\mathbb Z_2\), one has
\(\zeta(X_g,\sigma_g,E_g)=1\).  Hence an odd number of doublets has a finite
\(SU(2)\) anomaly.  This obstruction is not detected by the local cubic
descent calculation, because \(SU(2)\) has no independent symmetric cubic
invariant.
\end{proposition}

\begin{proof}
For one pseudoreal Weyl doublet the fermionic functional integral is not an
ordinary complex determinant function on the space of backgrounds; it is a
section of the real Pfaffian line associated with
\(D_{B,\sigma_M}^+\).  Equation~\eqref{eq:su2-global-anomaly-pfaffian} states
that parallel transport of this line around the background loop represented
by \((X_g,\sigma_g,E_g)\) multiplies the Pfaffian by
\((-1)^{\zeta(X_g,\sigma_g,E_g)}\).  This is the Dai--Freed holonomy
specialized to the pseudoreal \(SU(2)\) doublet, where the eta phase reduces
to the mod-two index sign.

For \(n\) independent doublets the Grassmann integral factors before the
fermions are coupled by any \(SU(2)\)-invariant mass term, and the anomaly
line is the tensor product of the \(n\) one-doublet Pfaffian lines.  Holonomy
is multiplicative under tensor product, so the sign is the \(n\)-th power of
the one-doublet sign:
\[
  \operatorname{Hol}(X_g,\sigma_g,E_g)
  =
  \bigl((-1)^{\zeta(X_g,\sigma_g,E_g)}\bigr)^n
  =
  (-1)^{n\,\zeta(X_g,\sigma_g,E_g)} .
\]
For the Witten mapping torus over \(S^4\), the mod-two index input gives
\(\zeta=1\).  The holonomy around the nontrivial loop is therefore
\((-1)^n\).

If the \(SU(2)\) is gauged, the fermion measure must descend from the space
of connections to the quotient by gauge transformations.  A loop in the
quotient whose lifted Pfaffian line has holonomy \(-1\) prevents such a
single-valued descent: after going around the loop, a putative gauge-invariant
partition function would have to equal its negative.  Thus odd \(n\) is
inconsistent with gauging this \(SU(2)\) symmetry.  If the same \(SU(2)\) is
kept as a background global symmetry, the same sign is an anomaly class that
must be reproduced by any infrared description.  Since the local cubic
invariant \(d^{abc}=\operatorname{tr}(t^a\{t^b,t^c\})\) vanishes for
\(\mathfrak{su}(2)\), this anomaly is a finite global obstruction rather than
an infinitesimal local descent anomaly.
\end{proof}

\section{Anomaly Inflow and Mapping Tori}

The finite anomaly in \eqref{eq:finite-background-anomaly} is represented by a
five-dimensional invertible response.  Its holonomy on a mapping torus gives
the finite transformation phase.  In the form needed here, an invertible
anomaly theory assigns a phase
\[
  \mathcal I_5[Y,\sigma_Y,B_Y]\in U(1)
\]
to every closed five-manifold \(Y\) equipped with a chosen spin structure
\(\sigma_Y\), global-symmetry background \(B_Y\), and any additional
geometric structures required by the fermions.  If
\((M,\sigma_M)=\partial(Y,\sigma_Y)\), the same theory assigns an anomaly line
over the space of boundary backgrounds, and the four-dimensional partition
function \(Z[M,\sigma_M,B]\) is a vector in that line.  Choosing a filling
\((Y,\sigma_Y,B_Y)\) trivializes the line by multiplying by the bulk phase.
Changing the filling changes the trivialization by the phase of the closed
structured five-manifold obtained by gluing the two fillings.

For a finite background gauge transformation \(g\), choose a path \(B_t\) from
\(B\) to \(B^g\) and identify the two ends by \(g\).  The resulting mapping
torus \(X_g\) carries the induced background \(B_g\).  The inflow statement is
the equality
\begin{equation}
  \frac{Z[M,\sigma_M,B^g]}{Z[M,\sigma_M,B]}
  =
  \mathcal I_5[X_g,\sigma_g,B_g],
  \label{eq:mapping-torus-inflow-phase}
\end{equation}
after a choice of local counterterm convention.  On the identity component of
the background gauge group, cutting the mapping torus open reduces
\eqref{eq:mapping-torus-inflow-phase} to the descent formula of the preceding
chapter.  Infinitesimally, the five-dimensional Chern--Simons response varies
by the boundary term \(-\int_M I_4^{(1)}\), while the boundary generating
functional varies by \(+\int_M I_4^{(1)}\).  For a disconnected component, the
mapping-torus phase can remain nontrivial even when the local polynomial
anomaly vanishes.

The \(SU(2)\) example above is exactly of this finite type.  Its anomaly
theory assigns
\begin{equation}
  \mathcal I^{SU(2)}_5[X_g,\sigma_g,E_g]
  =
  (-1)^{\zeta(X_g,\sigma_g,E_g)}
  \label{eq:su2-invertible-anomaly-theory}
\end{equation}
to the mapping torus with the induced rank-two bundle.  Equation
\eqref{eq:su2-global-anomaly-pfaffian} is therefore the mapping-torus inflow
law \eqref{eq:mapping-torus-inflow-phase}.  An odd number of fundamental Weyl
doublets has nontrivial \(\mathcal I^{SU(2)}_5\); an even number has the
trivial product phase for this test, and the Pfaffian line has no sign
holonomy around the generator detected by \(X_g\).

This language is also the precise version of the symmetry-protected
topological interpretation of anomalies.  A gapped short-range-entangled
five-dimensional phase with symmetry \(G_{\mathrm f}\) and no intrinsic
topological order has an invertible response \(\mathcal I_5\).  A
four-dimensional theory with anomaly \(\mathcal I_5\) can occur as a boundary
whose anomalous transformation is cancelled by the inverse bulk response.  In
mathematical treatments, deformation classes of fully extended
reflection-positive invertible responses with a specified tangential and
\(G_{\mathrm f}\)-background structure are often organized by characters of
the corresponding bordism groups, together with possible free Chern--Simons or
theta-type terms.  This cobordism organization is a classification framework
for invertible responses; each use in QFT still requires specifying the
background fields, spin or pin structure, locality assumptions, and the
normalization of the fermion determinant or Pfaffian line.

\section{Low-Degree Homotopy Groups in Four-Dimensional Applications}

The preceding example used \(\pi_4(SU(2))\) because a gauge transformation on
the compactified four-space \(S^4\) is a map \(S^4\to G\).  Other nearby
homotopy groups enter different constructions.  If a principal \(G\)-bundle on
\(S^4\) is described by transition data on the equator, the transition function
is a map \(S^3\to G\), so the instanton number is controlled by \(\pi_3(G)\)
and by the normalization of the corresponding characteristic class.  If a
four-dimensional nonlinear sigma model has target \(G\) and its Wess--Zumino
term is written by extending a field \(U:M^4\to G\) to a five-manifold, the
quantization is governed by the integral cohomology class of the closed
five-form used in the Wess--Zumino functional.  For the standard simple-group
case \(G=SU(N)\), \(N\ge3\), this quantization is tested on maps
\(S^5\to G\), hence by \(\pi_5(G)\simeq\mathbb Z\).  Thus
\(\pi_3,\pi_4,\pi_5\) play different roles and should not be conflated.

The following table records the low-degree groups most often met for compact
connected classical gauge groups.  The entries for \(SO(N)\) with \(N\ge7\),
\(U(N)\) with \(N\ge3\), \(SU(N)\) with \(N\ge3\), and \(Sp(N)\) with \(N\ge2\)
are the stable values in the displayed range, as given by Bott periodicity.
For \(d\ge2\), \(SO(N)\) and \(Spin(N)\) have the same \(\pi_d\); their
\(\pi_1\) differs because \(Spin(N)\to SO(N)\) is the universal double cover.
\[
\begin{array}{c|ccccc}
G & \pi_1(G) & \pi_2(G) & \pi_3(G) & \pi_4(G) & \pi_5(G) \\
\hline
U(1) & \mathbb Z & 0 & 0 & 0 & 0 \\
U(N),\ N\ge3 & \mathbb Z & 0 & \mathbb Z & 0 & \mathbb Z \\
SU(2)\simeq Sp(1) & 0 & 0 & \mathbb Z & \mathbb Z_2 & \mathbb Z_2 \\
SU(N),\ N\ge3 & 0 & 0 & \mathbb Z & 0 & \mathbb Z \\
SO(N),\ N\ge7 & \mathbb Z_2 & 0 & \mathbb Z & 0 & 0 \\
Spin(N),\ N\ge7 & 0 & 0 & \mathbb Z & 0 & 0 \\
Sp(N),\ N\ge2 & 0 & 0 & \mathbb Z & \mathbb Z_2 & \mathbb Z_2
\end{array}
\]
The low-rank orthogonal groups are exceptional because of accidental
isomorphisms:
\[
\begin{array}{c|ccc}
G & \pi_3(G) & \pi_4(G) & \pi_5(G) \\
\hline
Spin(3)\simeq SU(2) & \mathbb Z & \mathbb Z_2 & \mathbb Z_2 \\
Spin(4)\simeq SU(2)\times SU(2)
  & \mathbb Z^2 & \mathbb Z_2^2 & \mathbb Z_2^2 \\
Spin(5)\simeq Sp(2) & \mathbb Z & \mathbb Z_2 & \mathbb Z_2 \\
Spin(6)\simeq SU(4) & \mathbb Z & 0 & \mathbb Z .
\end{array}
\]
For simply connected exceptional compact groups
\[
  G=G_2,F_4,E_6,E_7,E_8
\]
one has
\[
  \pi_1(G)=\pi_2(G)=0,\qquad
  \pi_3(G)\simeq\mathbb Z,\qquad
  \pi_4(G)=\pi_5(G)=0 .
\]
There are further low-degree torsion groups in special cases, for example
\(\pi_6(SU(2))\simeq\mathbb Z_{12}\),
\(\pi_6(SU(3))\simeq\mathbb Z_6\), and
\(\pi_6(G_2)\simeq\mathbb Z_3\).  These become relevant in higher-dimensional
global-anomaly questions.

This table is a guide to the first topological obstructions, not a replacement
for the anomaly itself.  A modern finite anomaly is a phase of the fermion
determinant or Pfaffian line, and in its most invariant formulation it is an
invertible field theory on one higher dimension.  Homotopy groups detect
important test families of backgrounds; the complete torsion classification can
depend on spin structure, global form of the group, matter representation, and
bordism data.

\section{Vacuum Sectors and Locality}

Spontaneous symmetry breaking is a statement about phases of an infinite-volume
quantum system.  A useful comparison is a one-dimensional quantum mechanical
Hamiltonian
\[
  H=\frac{P^2}{2}+V(x),
  \qquad
  V(x)=V(-x),
\]
where \(V\) has two degenerate minima at \(x=\pm a\).
Figure~\ref{fig:double-well-finite-volume} uses this double-well model to fix
the finite-volume comparison: the exact ground state is symmetric, while
localized wave packets approximate broken-sector states only after the
infinite-volume limit is taken.
\begin{figure}[H]
  \centering
  \begin{tikzpicture}[>=Stealth,scale=0.92,baseline={(0,-0.1)}]
    \draw[->] (-2.7,0) -- (2.7,0) node[right] {\(x\)};
    \draw[->] (0,-0.15) -- (0,2.15) node[above] {\(V(x)\)};
    \draw[thick,domain=-2.25:2.25,smooth,variable=\x]
      plot ({\x},{0.18*(\x*\x-1.35)^2+0.18});
    \draw[thick,blue!70,domain=-1.85:-0.55,smooth,variable=\x]
      plot ({\x},{0.24+0.32*exp(-8*(\x+1.16)*(\x+1.16))});
    \draw[thick,blue!70,domain=0.55:1.85,smooth,variable=\x]
      plot ({\x},{0.24+0.32*exp(-8*(\x-1.16)*(\x-1.16))});
    \fill (-1.16,0.18) circle (1.6pt);
    \fill (1.16,0.18) circle (1.6pt);
    \node[below] at (-1.16,0) {\(-a\)};
    \node[below] at (1.16,0) {\(a\)};
    \node[blue!70] at (-1.72,0.64) {\(\ket{-a}\)};
    \node[blue!70] at (1.72,0.64) {\(\ket{+a}\)};
    \node at (1.9,1.65) {\(V(x)\)};
  \end{tikzpicture}
  \caption{A finite-dimensional double well has localized semiclassical states
  near the two minima, but the exact ground state is a symmetric
  superposition.}
  \label{fig:double-well-finite-volume}
\end{figure}
The localized semiclassical states \(\ket{+a}\) and \(\ket{-a}\) combine into
Hamiltonian eigenstates
\[
  \ket{\psi_\pm}
  =
  \frac1{\sqrt2}\bigl(\ket{+a}\pm\ket{-a}\bigr).
\]
The splitting between them is exponentially small in the Euclidean tunneling
action,
\[
  \Delta E\sim \ee^{-S_0},
  \qquad
  -\ddot x+V'(x)=0,
  \qquad
  x(\pm\infty)=\pm a .
\]
For a local quantum field theory with the analogous set of homogeneous
minima, the tunneling action is proportional to the spatial volume.  In the
infinite-volume limit the transition amplitude between distinct homogeneous
phases is suppressed as
\[
  \exp(-S_0\,\operatorname{Vol})
  \longrightarrow 0 .
\]
The pure vacua therefore define distinct phases rather than two nearly
degenerate levels of one finite-dimensional Hamiltonian.

Locality gives the corresponding superselection statement.  Let
\(\{\ket{\Omega_i}\}\) be translation-invariant pure vacua in infinite volume.
For any local operator \(\mathcal A(x)\), the vacuum matrix
\[
  A_{ij}:=\bra{\Omega_i}\mathcal A(0)\ket{\Omega_j}
\]
can be chosen diagonal in the phase label.  One way to see the constraint is
to compare two local operators \(\mathcal A(x)\) and \(\mathcal B(y)\) at
large spacelike separation.  Insert the vacuum sector decomposition and use
cluster decomposition in a pure phase:
\[
  \bra{\Omega_i}\mathcal A(x)\mathcal B(y)\ket{\Omega_j}
  =
  \sum_k A_{ik}B_{kj}+\hbox{terms vanishing as }|x-y|\to\infty .
\]
For spacelike separation, locality gives
\([\mathcal A(x),\mathcal B(y)]=0\), hence the vacuum matrices satisfy
\[
  \sum_k A_{ik}B_{kj}=\sum_k B_{ik}A_{kj}
\]
for all local \(\mathcal A,\mathcal B\).  The commuting family of vacuum
matrices is equivalently decomposed into its characters.  In a character
representation every element of the vacuum-sector algebra acts by a scalar.  If
a local observable had an off-diagonal matrix element between two distinct
characters, multiplication by another element of the algebra would give
different left and right scalar actions unless the two characters were equal.
Thus, in the pure-sector basis, local operators do not connect distinct phases:
\[
  \bra{\Omega_i}\mathcal A(x)\ket{\Omega_j}=0
  \qquad (i\ne j).
\]
This is superselection by locality and clustering, not a finite-volume level
splitting statement.

\section{Spontaneous Symmetry Breaking}

\begin{definition}[Internal spontaneous symmetry breaking datum]
\label{def:internal-ssb-datum}
Consider a Lorentz-invariant quantum theory in infinite spatial volume with an
exact continuous internal symmetry group \(G\).  Choose one pure phase and let
\(\ket{\vac}\) be its vacuum vector.  The unbroken subgroup is
\[
  H
  =
  \{h\in G:U(h)\ket{\vac}=\ket{\vac}\}.
\]
The symmetry is spontaneously broken from \(G\) to \(H\) when \(H\) is a proper
subgroup.  Let \(\mathfrak g=\operatorname{Lie}(G)\) and
\(\mathfrak h=\operatorname{Lie}(H)\).  Choose an \(H\)-stable complement
\(\mathfrak m\) so that
\[
  \mathfrak g=\mathfrak h\oplus\mathfrak m,
  \qquad
  [\mathfrak h,\mathfrak m]\subset\mathfrak m .
\]
Elements \(T_i\in\mathfrak h\) generate unbroken transformations and elements
\(X_a\in\mathfrak m\) generate broken directions.
\end{definition}

This is an infinite-volume datum, not an automatic consequence of writing a
continuous symmetry into the Lagrangian.  In low dimensions the infrared
fluctuations recorded in
Section~\ref{sec:low-dimensional-continuous-ssb-no-go} can prevent the
nonzero order parameter required by this definition.

\begin{definition}[Goldstone effective action in CCWZ coordinates]
\label{def:goldstone-effective-action}
Goldstone fields are local coordinates on the coset \(G/H\).  A coset
representative may be written
\begin{equation}
  \xi(\pi(x))
  =
  \exp\left\{\frac{\ii}{f}\pi^a(x)X_a\right\}.
  \label{eq:goldstone-coset-coordinate}
\end{equation}
For \(g\in G\), the defining nonlinear action is
\begin{equation}
  g\,\xi(\pi)
  =
  \xi(\pi')\,h(g,\pi),
  \qquad
  h(g,\pi)\in H .
  \label{eq:ccwz-nonlinear-action}
\end{equation}
The compensator \(h(g,\pi)\) is determined by the chosen coset section and by
the requirement that \(g\xi(\pi)\) be written again in the form
\(\xi(\pi')H\).

The Maurer--Cartan form decomposes as
\begin{equation}
  \xi^{-1}\dd\xi
  =
  \ii e^a X_a+\ii\omega^i T_i .
  \label{eq:ccwz-maurer-cartan}
\end{equation}
The one-forms \(e^a=e^a_{\mu}\dd x^\mu\) are the covariant vielbein on the
Goldstone target \(G/H\), while \(\omega^i\) is an \(H\)-connection.  If
\(G/H\) carries an invariant metric \(G_{ab}\) on the broken tangent space, the
leading Goldstone action is
\begin{equation}
  S_{\mathrm{Goldstone}}
  =
  \frac{f^2}{2}
  \int G_{ab}\,
  e^a_\mu e^{b\mu}\,\dd^4x
  +\cdots .
  \label{eq:goldstone-leading-action}
\end{equation}
The ellipsis denotes higher-derivative terms and explicit symmetry-breaking
terms.  No potential appears for exact Goldstone coordinates because constant
maps into \(G/H\) label symmetry-related vacua.
\end{definition}

\section{Goldstone Theorem as a Spectral Statement}

\begin{proposition}[Goldstone spectral pole from a broken current]
\label{prop:goldstone-spectral-pole}
The spectral input behind \eqref{eq:goldstone-leading-action} is as follows.
Assume a conserved current \(j_A^\mu\), a local operator \(\mathcal O\), and a
nonzero order parameter
\begin{equation}
  \langle\vac,[Q_A,\mathcal O(0)]\vac\rangle
  =
  \langle\vac,\delta_A\mathcal O(0)\vac\rangle
  \ne0 ,
  \label{eq:goldstone-order-parameter}
\end{equation}
where \(Q_A\) is obtained as the infinite-volume limit of smeared charges.
Assume also that local contact ambiguities in the time-ordered product have
been fixed so that the nonlocal part of the mixed current-operator distribution
has the Lorentz-covariant scalar spectral form near \(p=0\): modulo a local
polynomial in \(p\),
\[
  \widehat G_A^\mu(p)_{\rm nonloc}
  =
  p^\mu F_A(p^2)
  +
  \widehat G_{A,\perp}^\mu(p),
  \qquad
  p_\mu\widehat G_{A,\perp}^\mu(p)=0,
\]
where \(F_A\) is a scalar boundary-value distribution with a
K\"all\'en--Lehmann representation in the broken channel.
The Ward identity for the time-ordered distribution is
\begin{equation}
  \partial_\mu
  \langle\vac,T\{j_A^\mu(x)\mathcal O(0)\}\vac\rangle
  =
  -\ii\delta^{(4)}(x)
  \langle\vac,\delta_A\mathcal O(0)\vac\rangle .
  \label{eq:goldstone-ward-distribution}
\end{equation}
Fourier transforming gives a distribution \(\widehat G_A^\mu(p)\) satisfying
\[
  p_\mu\widehat G_A^\mu(p)
  =
  \langle\vac,\delta_A\mathcal O(0)\vac\rangle .
\]
Since the right-hand side is nonzero at \(p=0\), \(F_A(p^2)\) must contain a
\(1/(p^2-\ii0)\)-type massless singularity.  In the K\"all\'en--Lehmann
language this singularity is a massless contribution to the
current-operator spectral measure.
\end{proposition}

\begin{proof}
Let \(c_A=\langle\vac,\delta_A\mathcal O(0)\vac\rangle\neq0\).
Integrating \eqref{eq:goldstone-ward-distribution} against a test function
equal to one near the origin gives the broken-charge matrix element
\eqref{eq:goldstone-order-parameter}; Fourier transformation converts the
distributional divergence into multiplication by \(p_\mu\).  After the chosen
local contact polynomial has been separated, the transverse part drops out and
the Ward identity reads, for the nonlocal scalar channel,
\[
  p^2F_A(p^2)+p_\mu P_A^\mu(p)=c_A ,
\]
where \(P_A^\mu(p)\) is the remaining local polynomial vector contribution.
The contraction \(p_\mu P_A^\mu(p)\) has no constant term at \(p=0\).  If
\(F_A\) were regular at \(p^2=0\), then \(p^2F_A(p^2)\) would also vanish at
the origin as a boundary-value distribution modulo higher-order local
polynomial terms, contradicting \(c_A\neq0\).  Thus the nonlocal scalar channel
contains
\[
  F_A(p^2)=\frac{c_A}{p^2-\ii0}+\text{less singular terms}
\]
up to the sign fixed by the Fourier convention.  In a scalar spectral
representation
\[
  F_A(p^2)=
  \int_{0}^{\infty}\frac{\dd\nu_A(\mu^2)}{p^2+\mu^2-\ii0}
  +\text{local polynomial},
\]
such an inverse \(p^2\) boundary value is precisely an atom of \(\dd\nu_A\) at
\(\mu^2=0\).  This is the massless Goldstone contribution to the spectral
measure.
\end{proof}

Equivalently, for each broken direction there are zero-mass one-particle
states \(\ket{\pi_a(p)}\) created by the current:
\begin{equation}
  \bra{\vac}j_A^\mu(0)\ket{\pi_a(p)}
  =
  \ii F_{Aa}p^\mu,
  \qquad p^2=0 .
  \label{eq:goldstone-current-matrix-element}
\end{equation}
If the translation generators have spectrum contained in the closed forward
light cone and the vacuum is the unique translation-invariant vector in the
chosen phase, the matrix \(F_{Aa}\) has rank equal to the number of broken
generators.  These massless states are the quanta described by the Goldstone
coordinates \(\pi^a(x)\).  This one-particle-per-broken-generator conclusion
is the zero commutator-density special case of the non-Lorentz counting
theorem in Section~\ref{sec:nonlorentz-goldstone-counting}.

\begin{remark}[Goldstone theorem and gauged directions]
\label{rem:goldstone-gauge-brst-higgs}
The spectral argument above is a theorem about a global symmetry current acting
on the physical positive Hilbert space.  It assumes that \(Q_A\) is a charge
defined on physical states and that \(\delta_A\mathcal O\) is the variation of
a gauge-invariant local observable or of a local field in an ordinary
positive-metric representation.

Gauge redundancy is different.  In a covariant gauge-fixed description the
state space before the physical quotient is a Krein space, as in
Chapter~\ref{ch:volume-ii-18}, and physical states are BRST cohomology
classes.  A scalar coordinate that moves along a gauged orbit is then a
would-be Goldstone coordinate, not by itself a physical massless particle.  At
the level of gauge theories with scalar matter, this is the
Brout--Englert--Higgs mechanism developed in
Section~\ref{sec:beh-mechanism}.  At
quadratic order in the Abelian Higgs model, using here the conventional
explicit-coupling coordinate \(D_\mu=\partial_\mu-\ii gA_\mu\), write
\[
  \phi(x)=\frac{1}{\sqrt2}\bigl(v+h(x)+\ii\pi(x)\bigr),
  \qquad
  m_A=gv .
\]
The scalar kinetic term contains
\begin{equation}
  |D_\mu\phi|^2
  =
  \frac12(\partial_\mu h)^2
  +
  \frac12(\partial_\mu\pi-m_A A_\mu)^2
  +\text{higher powers of }h,\pi,A .
  \label{eq:abelian-higgs-quadratic-mixing}
\end{equation}
Equivalently, away from zeros of \(v+h\), the gauge-invariant vector
\[
  \mathcal A_\mu
  =
  A_\mu-\frac{1}{m_A}\partial_\mu\pi
\]
has the Proca mass term \(\frac12m_A^2\mathcal A_\mu\mathcal A^\mu\) at
quadratic order.  In an \(R_\xi\) gauge one instead uses the gauge-fixing
functional
\[
  F_\xi=\partial^\mu A_\mu+\xi m_A\pi ,
\]
which cancels the \(A_\mu\partial^\mu\pi\) mixing and assigns the unphysical
fields \(\pi\) and the Faddeev--Popov ghosts the gauge-dependent mass
\(\sqrt{\xi}\,m_A\).  The longitudinal gauge polarization, the would-be
Goldstone coordinate, and the ghost-antighost variables form BRST doublet or
quartet sectors, so they have no independent positive-norm particle states in
the BRST cohomology.  The physical spectrum contains a massive vector particle
with three polarizations and the gauge-invariant scalar excitation \(h\), plus
any Goldstone particles associated with broken global symmetries that remain
after the gauge redundancy has been quotiented.
\end{remark}

\section{Non-Lorentz Goldstone Counting and Type-B Modes}
\label{sec:nonlorentz-goldstone-counting}

Lorentz invariance was load-bearing in the preceding spectral rank statement.
At finite density, in a magnet, or in a microscopic Hamiltonian system with no
boost symmetry, a broken internal generator still produces low-energy
spectral weight, but several broken generators can supply only one canonical
degree of freedom.  The missing datum is the density of charge commutators in
the selected thermodynamic phase.

\begin{definition}[Commutator-density matrix]
\label{def:thermodynamic-commutator-density-matrix}
Let \(\omega\) be a translation-invariant pure or extremal phase of a local
Hamiltonian system in infinite spatial volume, and let
\(j_a^\mu=(j_a^0,\mathbf j_a)\) be conserved currents for an exact internal
symmetry algebra.  Choose smooth charge cutoffs \(f_R(x)\) which equal one on
a ball of radius \(R\), vanish outside a slightly larger ball, and have
boundary gradients supported only in the collar.  Write
\[
  Q_a(R)=\int\dd^d x\,f_R(x)j_a^0(t,x),
  \qquad
  |B_R|=\int\dd^d x\,\mathbf 1_{|x|<R}.
\]
The commutator-density matrix of the broken charges is
\begin{equation}
  \rho_{ab}
  =
  -\ii\lim_{R\to\infty}
  \frac{1}{|B_R|}
  \omega\!\left([Q_a(R),Q_b(R)]\right),
  \label{eq:goldstone-commutator-density-matrix}
\end{equation}
when the limit exists and is independent of the admissible cutoff family.
If the regulated local algebra has no central extension and the charges obey
\[
  [Q_a,Q_b]=\ii f_{ab}{}^c Q_c,
\]
and if the charge densities
\[
  n_c=\lim_{R\to\infty}|B_R|^{-1}\omega(Q_c(R))
\]
exist, then
\[
  \rho_{ab}=f_{ab}{}^c n_c .
\]
The definition is thermodynamic: it assumes locality good enough to remove
the smearing collar, finite charge densities, and a phase in which the
large-volume current commutators have a stable limit.
\end{definition}

\begin{quotedtheorem}[Internal Nambu--Goldstone counting without Lorentz invariance]
\label{qthm:nonlorentz-ng-counting}
\emph{Proof reference.}\footnote{The inequality in this language goes back to
Nielsen and Chadha \cite{NielsenChadha1976GoldstoneCounting}.  The
charge-density rank formula and its effective-action interpretation are due
to Watanabe--Brauner \cite{WatanabeBrauner2011NGBChargeDensity} and
Watanabe--Murayama
\cite{WatanabeMurayama2012UnifiedNGB,WatanabeMurayama2014NonrelativisticEL};
Leutwyler's nonrelativistic effective Lagrangian analysis is the earlier EFT
source for the same first-order term in magnets
\cite{Leutwyler1994NonrelativisticEL}.}
Let an internal continuous symmetry be spontaneously broken from \(G\) to
\(H\) in a translation-invariant thermodynamic phase.  Assume local conserved
current densities, a well-defined infinite-volume broken-charge action on
local operators, finite charge densities, clustering or regularity sufficient
to define the matrix \(\rho_{ab}\) in
\eqref{eq:goldstone-commutator-density-matrix}, and no additional gapless
singularities that invalidate the low-energy derivative expansion.  Let
\[
  n_{\rm BS}=\dim(G/H)
\]
be the number of broken internal generators and let the antisymmetric matrix
\(\rho\) be restricted to those broken directions.  Then
\begin{equation}
  n_{\rm type\text{-}B}
  =
  \frac12\operatorname{rank}\rho,
  \qquad
  n_{\rm type\text{-}A}
  =
  n_{\rm BS}-\operatorname{rank}\rho,
  \qquad
  n_{\rm NG}
  =
  n_{\rm BS}-\frac12\operatorname{rank}\rho .
  \label{eq:watanabe-murayama-counting}
\end{equation}
Here type-B modes are the modes paired by the nonzero commutator-density
blocks, while type-A modes are the unpaired broken directions.  The
Nielsen--Chadha count \(n_{\rm type\text{-}A}+2n_{\rm type\text{-}B}\ge
n_{\rm BS}\) is saturated under the stated regularity hypotheses.  In a
Lorentz-invariant vacuum, or more generally whenever all broken charge
commutator densities vanish, \(\rho=0\) and the formula reduces to one
Nambu--Goldstone mode for each broken internal generator.
\end{quotedtheorem}

The mechanism behind the rank formula is finite-dimensional before it is a
many-body theorem.  Let \(\pi^a\) be local coordinates on the broken orbit.
The most general leading real-time effective Lagrangian may contain a term
first order in time derivatives,
\begin{equation}
  L_{\rm time}
  =
  c_a(\pi)\dot\pi^a
  +
  \frac12 g_{ab}(\pi)\dot\pi^a\dot\pi^b+\cdots ,
  \qquad
  \partial_a c_b-\partial_b c_a
  =
  \rho_{ab}+O(\pi).
  \label{eq:nonlorentz-goldstone-first-order-form}
\end{equation}
By a linear change of broken generators, the constant antisymmetric matrix
\(\rho\) is brought to Darboux blocks
\[
  \rho
  =
  \bigoplus_{\alpha=1}^{r}
  \begin{pmatrix}0&\lambda_\alpha\\-\lambda_\alpha&0\end{pmatrix}
  \oplus 0,
  \qquad
  r=\frac12\operatorname{rank}\rho .
\]
Each block supplies a canonical pair.  In a block with coordinates
\((q_\alpha,p_\alpha)\), the term
\(\frac{\lambda_\alpha}{2}(q_\alpha\dot p_\alpha-p_\alpha\dot q_\alpha)\)
already gives the symplectic form.  The two broken coordinates are therefore
one phase-space degree of freedom, not two independent configuration-space
waves.  Directions in the kernel of \(\rho\) have no such first-order
symplectic pairing and require the ordinary second-order kinetic term in
\eqref{eq:nonlorentz-goldstone-first-order-form}; they count one mode per
coordinate.

The common dispersion laws come from adding the leading spatial stiffness,
not from the rank formula alone.  For a type-A coordinate,
\[
  L_A=\frac12 G\dot\varphi^2-\frac12 K(\nabla\varphi)^2
  \quad\Longrightarrow\quad
  \omega^2=\frac{K}{G}|\mathbf k|^2 .
\]
For one type-B block,
\[
  L_B=
  \frac{\lambda}{2}(q\dot p-p\dot q)
  -\frac{K}{2}\bigl((\nabla q)^2+(\nabla p)^2\bigr),
\]
the equations of motion give
\begin{equation}
  \lambda\dot q=K\nabla^2p,
  \qquad
  \lambda\dot p=-K\nabla^2q,
  \qquad
  \omega=\frac{K}{|\lambda|}|\mathbf k|^2 .
  \label{eq:type-b-quadratic-dispersion}
\end{equation}
Thus type-A modes are commonly linear and type-B modes commonly quadratic.
The word ``commonly'' matters: long-range interactions, dissipative open-system
terms, Lifshitz scaling, or a vanishing stiffness can change the power law.
The counting theorem counts gapless branches once a broken internal-symmetry
phase and its regular low-energy expansion exist; it is not by itself a
dispersion theorem.

Two physical laboratories keep the theorem honest.  First consider a
finite-density relativistic complex scalar or superfluid.  In a grand
canonical description one fixes the unbroken time evolution
\(H_\mu=H-\mu Q\), and the \(U(1)\) symmetry is broken by the condensate.  The
phase field has an effective action
\[
  L_{\rm sf}^{(2)}
  =
  \frac12\chi\,\dot\theta^2
  -
  \frac12\rho_s(\nabla\theta)^2+\cdots ,
  \qquad
  \chi=\frac{\partial n}{\partial\mu}>0,\quad \rho_s>0 .
\]
There is one broken internal generator, hence \(\rho=0\), and
\[
  \omega_{\rm sf}(\mathbf k)
  =
  \sqrt{\rho_s/\chi}\,|\mathbf k|+O(|\mathbf k|^3).
\]
The chemical potential has broken boost invariance and selected a rest frame,
but it has not created a nonzero commutator density among two broken internal
charges, so the phonon remains type-A.

Now compare a Heisenberg ferromagnet in \(d\ge3\) at zero temperature, where
the low-dimensional obstruction below does not remove the ordered phase.  The
microscopic symmetry is \(SU(2)\), the magnetization density
\[
  m=\lim_{V\to\infty}\frac{1}{V}\omega(Q_3)
\]
selects the unbroken \(U(1)\) generated by \(Q_3\), and the broken charges are
\(Q_1,Q_2\).  Since
\([Q_1,Q_2]=\ii Q_3\),
\[
  \rho_{12}=m,\qquad
  \rho=
  \begin{pmatrix}0&m\\-m&0\end{pmatrix}.
\]
The rank is two, so \(n_{\rm BS}=2\) gives
\[
  n_{\rm type\text{-}B}=1,\qquad
  n_{\rm type\text{-}A}=0,\qquad
  n_{\rm NG}=1 .
\]
The spin-coherent effective action for a unit vector
\(\mathbf n=(\pi_1,\pi_2,\sqrt{1-\pi^2})\) has, to quadratic order,
\[
  L_{\rm F}^{(2)}
  =
  \frac{m}{2}(\pi_1\dot\pi_2-\pi_2\dot\pi_1)
  -
  \frac{A}{2}\bigl((\nabla\pi_1)^2+(\nabla\pi_2)^2\bigr),
\]
and therefore
\[
  \omega_{\rm F}(\mathbf k)=\frac{A}{|m|}|\mathbf k|^2+O(|\mathbf k|^4).
\]
Two broken spin rotations have become one quadratically dispersing magnon
because the background magnetization is the expectation value of their
commutator.

The antiferromagnet is the useful contrast.  It can have the same internal
breaking pattern \(SU(2)\to U(1)\) but zero uniform spin density, hence
\(\rho=0\).  Its leading staggered-order action is second order in time,
\[
  L_{\rm AF}^{(2)}
  =
  \frac12\chi_\perp(\dot\pi_1^2+\dot\pi_2^2)
  -
  \frac12\rho_s\bigl((\nabla\pi_1)^2+(\nabla\pi_2)^2\bigr),
\]
so the two broken generators give two linear spin waves rather than one
quadratic magnon.  The difference is not the coset \(SU(2)/U(1)\), which is
the same in the two examples; it is the phase's charge-density matrix.

Finally, the count above is only for internal symmetries.  At finite density
one normally works with the unbroken Hamiltonian \(H_\mu\), and boosts may be
broken by the state; in solids, translations and rotations can also be broken.
Broken spacetime generators act on local fields with explicit coordinate or
derivative pieces, and inverse-Higgs constraints can make their putative
Goldstone coordinates redundant.  Those spacetime-redundancy questions are
not entries of \(n_{\rm BS}\) in
\eqref{eq:watanabe-murayama-counting}.  Likewise,
Section~\ref{sec:low-dimensional-continuous-ssb-no-go} constrains whether the
ordered thermodynamic phase exists in the first place.  Once the no-go theorem
has removed the order parameter, there is no broken phase for
\eqref{eq:watanabe-murayama-counting} to count.

\section{Low-Dimensional Continuous-Symmetry No-Go Theorems}
\label{sec:low-dimensional-continuous-ssb-no-go}

The Goldstone proposition is conditional: if a conserved global current acts
on a positive physical Hilbert space and if a local order parameter has a
nonzero variation, then massless spectral weight follows.  The low-dimensional
no-go theorems remove the premise in important settings.  They do not say that
all two-dimensional systems are disordered, and they do not contradict
Goldstone's theorem; they say that ordinary local long-range order for a
continuous internal symmetry is destroyed by infrared fluctuations under
specific hypotheses.

\begin{quotedtheorem}[Coleman no-Goldstone theorem in \(1+1\) dimensions]
\label{qthm:coleman-no-goldstone-two-dimensions}
\emph{Proof reference.}\footnote{See
Coleman's original positive-metric Wightman proof
\cite{Coleman1973NoGoldstone}.}
Let a relativistic \(1+1\)-dimensional quantum field theory satisfy the
positive-metric Wightman assumptions, translation invariance, locality, and the
spectrum condition.  Let a continuous internal symmetry be implemented by a
conserved local current \(j_A^\mu\) whose compactly supported charge
approximants generate the local field variations.  The order field
\(\mathcal O\) is required to be an ordinary Wightman field or local observable:
its compactly smeared vectors have finite Hilbert norm, its correlation
functions are tempered distributions, and the current--field commutator is a
local distribution generating \(\delta_A\mathcal O\).  Then there is no such
\(\mathcal O\) for which
\[
  \langle\vac,\delta_A\mathcal O(0)\vac\rangle\ne0
\]
under these hypotheses.  Equivalently, an ordinary relativistic
\(1+1\)-dimensional positive-metric local QFT has no Goldstone boson produced
by spontaneous breaking of a continuous internal symmetry.
\end{quotedtheorem}

Clustering is not being quoted as an original Coleman hypothesis here.  It is
a separate phase-selection condition often imposed when one wants to distinguish
an extremal vacuum sector from a mixture of sectors.  Coleman's contradiction
uses the positive metric, locality, current Ward identity, spectral support,
and the Wightman infrared regularity of the local order field.  If a later
formulation replaces part of that infrared regularity discussion by a
cluster-property assumption, that is a separate theorem statement with its own
domain; the present chapter does not use it as the proof reference for
Theorem~\ref{qthm:coleman-no-goldstone-two-dimensions}.

Here is the infrared mechanism in the form needed for the present chapter.
Suppose the order-parameter variation in
\eqref{eq:goldstone-order-parameter} were nonzero in \(1+1\) dimensions.
The same Ward identity that led to
Proposition~\ref{prop:goldstone-spectral-pole} would force a longitudinal
massless contribution in the mixed current-operator distribution.  Positivity
then lets one compare the required longitudinal channel with the
infrared-regulated massless scalar covariance.  At equal time, with spatial
momenta cut off by \(\Lambda\) and with an infrared mass \(m\),
\begin{equation}
  \int_{-\Lambda}^{\Lambda}
  \frac{\dd k}{2\pi}\,
  \frac{1}{2\sqrt{k^2+m^2}}
  =
  \frac{1}{2\pi}\operatorname{arsinh}\frac{\Lambda}{m}
  =
  \frac{1}{2\pi}\log\frac{2\Lambda}{m}+o(1)
  \qquad(m\downarrow0).
  \label{eq:coleman-massless-scalar-log}
\end{equation}
The logarithm is the obstruction.  A compactly supported smeared local field
must have finite vacuum norm.  The would-be Goldstone coordinate in two
spacetime dimensions instead behaves like a massless scalar whose zero-mode
fluctuations grow without bound.  The sharp Coleman proof does not assume the
field is literally free; it uses smeared current Ward identities, locality,
the spectral condition, and positive Hilbert-space norms to force the same
massless longitudinal infrared singularity and then shows that it is
incompatible with a local charged order parameter.

This distinction is essential in bosonized two-dimensional models.  Derivative
currents, compact-boson vertex operators with a specified radius, and
topological or winding sectors can be perfectly good local or semi-local QFT
data.  What Coleman forbids is the unsmeared massless scalar expectation, or
an ordinary local field whose vacuum expectation would choose a point on a
continuous internal orbit.  Sine-Gordon soliton sectors, Thirring currents,
WZW currents, and compact torus CFT vertex operators therefore do not by
themselves constitute continuous-symmetry breaking.

\begin{quotedtheorem}[Mermin--Wagner theorem for isotropic Heisenberg magnets]
\label{qthm:mermin-wagner-heisenberg}
\emph{Proof reference.}\footnote{This is the nearest-neighbor quantum-spin
model class of Mermin and Wagner \cite{MerminWagner1966}.  The statement below
is intentionally narrower than later Mermin--Wagner-type theorems.}
Let \(\Lambda_L\subset\mathbb Z^d\), \(d=1,2\), be a periodic box, and let the
local algebra be
\(\mathcal A_{\Lambda_L}=\bigotimes_{x\in\Lambda_L}M_{2S+1}(\mathbb C)\) with
spin-\(S\) generators \(\mathbf S_x\).  The global \(SU(2)\) symmetry acts by
the diagonal spin rotation.  For the ferromagnet take
\[
  H_{L,h}^{\rm F}
  =
  -J\sum_{\langle x,y\rangle}\mathbf S_x\cdot\mathbf S_y
  -h\sum_x S_x^z,
  \qquad J>0,\ h>0,
\]
and for the bipartite antiferromagnet replace the first sign by \(+J\) and
the field by a staggered field
\(-h\sum_x(-1)^{|x|}S_x^z\).  At any finite inverse temperature
\(\beta<\infty\), form the finite-volume Gibbs state, take the thermodynamic
limit \(L\to\infty\) along periodic boxes, and only then send the selecting
field \(h\downarrow0\).  The ferromagnetic magnetization and the
antiferromagnetic staggered magnetization vanish in this order of limits.
Nearest-neighbor coupling is the declared interaction regularity; equivalently
the exchange has a finite second moment, and the small-momentum denominator in
the Bogoliubov estimate is quadratic.
\end{quotedtheorem}

No clustering assumption appears in this quoted spin theorem.  The conclusion
is obtained in the external-field thermodynamic-limit prescription and applies
to any limit point reached that way.  If one later imposes extremality or
clustering to select a pure Gibbs/KMS phase, that extra phase-selection
condition does not replace the Bogoliubov infrared estimate.

\begin{quotedtheorem}[Hohenberg theorem for low-dimensional Bose condensation]
\label{qthm:hohenberg-bose-no-odlro}
\emph{Proof reference.}\footnote{See Hohenberg's continuum equilibrium
Bogoliubov-inequality argument \cite{Hohenberg1967}.}
Consider a nonrelativistic Bose field in a box
\(\Lambda_L\subset\mathbb R^d\), \(d=1,2\), at fixed finite density and finite
inverse temperature \(\beta<\infty\).  The local field algebra is generated by
smeared Bose fields \(\psi(f)\), the internal symmetry is the global
\(U(1)\) action \(\psi\mapsto e^{i\alpha}\psi\), and the Hamiltonian has the
standard kinetic energy plus a stable tempered two-body interaction for which
the finite-density Gibbs thermodynamic limit, finite kinetic-energy density,
and the \(f\)-sum/current-commutator bound used in the proof exist.  A concrete
sufficient short-range hypothesis for the interaction part is
\[
  \int_{\mathbb R^d}(1+|x|^2)|V(x)|\,\dd^d x<\infty
\]
together with the usual stability lower bound.  Introduce a uniform
symmetry-breaking source for \(\psi+\psi^\dagger\), take \(L\to\infty\) at
fixed source, and then send the source to zero.  In that order of limits the
condensate amplitude vanishes; equivalently the ordinary off-diagonal
long-range-order parameter
\[
  \lim_{|x|\to\infty}\langle\psi^\dagger(x)\psi(0)\rangle
\]
is zero whenever this limit is the ODLRO diagnostic under consideration.
\end{quotedtheorem}

Clustering is again not an input to the quoted Hohenberg theorem.  The theorem
removes the ODLRO order parameter itself.  In a clustering state this would
also force the factorized condensate amplitude to vanish, but the absence of
clustering is not an escape from the infrared bound.

The spin-wave calculation is physical motivation, not a proof of either
quoted theorem.  If a continuous order parameter of stiffness \(\rho_s\)
existed at temperature \(T=\beta^{-1}\), its angular fluctuations would contain
\begin{equation}
  \langle \pi^2\rangle_{\rm IR}
  \sim
  \frac{T}{\rho_s}
  \int_{|k|<\Lambda}
  \frac{\dd^d k}{(2\pi)^d}\,
  \frac{1}{k^2+m^2},
  \label{eq:mwh-spin-wave-integral}
\end{equation}
where \(m\) is an infrared field or inverse box-size regulator.  The
dimension split is
\begin{align}
  d=1:\quad&
  \int_{-\Lambda}^{\Lambda}
  \frac{\dd k}{2\pi}\,\frac1{k^2+m^2}
  =
  \frac{1}{\pi m}\arctan\frac{\Lambda}{m}
  \sim \frac{1}{2m},
  \nonumber\\
  d=2:\quad&
  \int_{|k|<\Lambda}
  \frac{\dd^2 k}{(2\pi)^2}\,\frac1{k^2+m^2}
  =
  \frac{1}{4\pi}\log\frac{\Lambda^2+m^2}{m^2}
  =
  \frac{1}{2\pi}\log\frac{\Lambda}{m}+o(1),
  \nonumber\\
  d=3:\quad&
  \int_{|k|<\Lambda}
  \frac{\dd^3 k}{(2\pi)^3}\,\frac1{k^2+m^2}
  =
  \frac{1}{2\pi^2}
  \left(\Lambda-m\arctan\frac{\Lambda}{m}\right)
  \longrightarrow \frac{\Lambda}{2\pi^2}.
  \label{eq:mwh-infrared-dimension-split}
\end{align}
Thus the long-wavelength variance diverges in \(d=1,2\) and is infrared-finite
in \(d=3\).  This Gaussian estimate explains the mechanism but does not supply
the Gibbs/KMS construction, positivity inequality, order of limits, or
operator algebra needed for the theorem statements.

For the nearest-neighbor Heisenberg ferromagnet the relevant rigorous
mechanism can be seen directly.

\begin{proposition}[Bogoliubov bound in the nearest-neighbor Heisenberg model]
\label{prop:heisenberg-bogoliubov-bound}
In the finite periodic ferromagnet of
Theorem~\ref{qthm:mermin-wagner-heisenberg}, set
\[
  m_{L,h}=\frac1{|\Lambda_L|}\sum_{x\in\Lambda_L}
  \langle S_x^z\rangle_{L,h}
\]
and use normalized Fourier modes
\[
  \widehat S_k^a
  =
  |\Lambda_L|^{-1/2}
  \sum_{x\in\Lambda_L}e^{ik\cdot x}S_x^a,
  \qquad
  k\in\operatorname{BZ}_L .
\]
There is a spin-dependent constant \(c_S\) such that, for every \(h>0\),
\begin{equation}
  m_{L,h}^2
  \le
  \beta S(S+1)
  \left[
    \frac1{|\Lambda_L|}
    \sum_{k\in\operatorname{BZ}_L\setminus\{0\}}
    \frac{1}{h\,m_{L,h}+c_SJ\,\epsilon(k)}
  \right]^{-1},
  \qquad
  \epsilon(k)=2\sum_{j=1}^d(1-\cos k_j).
  \label{eq:heisenberg-bogoliubov-finite-volume-bound}
\end{equation}
Consequently the field-selected spontaneous magnetization vanishes in
\(d=1,2\).
\end{proposition}

\begin{proof}
Use the finite-temperature Bogoliubov inequality
\[
  \langle\{A,A^\dagger\}\rangle_{L,h}
  \ge
  \frac{2}{\beta}\,
  \frac{|\langle[C,A]\rangle_{L,h}|^2}
       {\langle[[C,H_{L,h}^{\rm F}],C^\dagger]\rangle_{L,h}}
\]
with \(A=\widehat S_{-k}^x\) and \(C=\widehat S_k^y\).  The commutator is
\[
  \langle[\widehat S_k^y,\widehat S_{-k}^x]\rangle_{L,h}
  =
  -i\,m_{L,h}.
\]
The external field contributes \(h\,m_{L,h}\) to the normalized double
commutator.  A direct nearest-neighbor spin-commutator calculation gives the
exchange bound
\[
  \langle[[\widehat S_k^y,-J\sum_{\langle x,y\rangle}
  \mathbf S_x\cdot\mathbf S_y],\widehat S_{-k}^y]\rangle_{L,h}
  \le
  c_SJ\,\epsilon(k),
\]
where the local spin bound \(\|\mathbf S_x\|^2\le S(S+1)\) absorbs the harmless
constant.  Thus
\begin{equation}
  \langle\{\widehat S_{-k}^x,\widehat S_k^x\}\rangle_{L,h}
  \ge
  \frac{2m_{L,h}^2}
       {\beta\,[h\,m_{L,h}+c_SJ\,\epsilon(k)]}.
  \label{eq:heisenberg-mode-bogoliubov-lower-bound}
\end{equation}
Summing this inequality over nonzero \(k\) and using Parseval,
\[
  \frac1{|\Lambda_L|}\sum_{k\in\operatorname{BZ}_L}
  \langle\{\widehat S_{-k}^x,\widehat S_k^x\}\rangle_{L,h}
  =
  \frac1{|\Lambda_L|}\sum_x
  \langle\{S_x^x,S_x^x\}\rangle_{L,h}
  \le
  2S(S+1),
\]
gives \eqref{eq:heisenberg-bogoliubov-finite-volume-bound}, after changing
the inessential numerical constant.  Finally
\(\epsilon(k)\le |k|^2\) for small \(k\).  If
\(\lim_{h\downarrow0}\limsup_{L\to\infty}m_{L,h}\) were positive, then the
thermodynamic-limit version of the denominator would dominate
\[
  \int_{|k|<\Lambda}\frac{\dd^dk}{(2\pi)^d}\,
  \frac{1}{h\,m+c|k|^2},
\]
which diverges as \(h\downarrow0\) in \(d=1,2\).  The bound then forces
\(m=0\), a contradiction.  The staggered antiferromagnetic estimate uses the
same inequality with the staggered Fourier component and the ordering wave
vector shifted by the bipartite wave vector.
\end{proof}

The model-specific qualifications are part of the statement, not marginal
notes.  Finite volumes have no spontaneous order parameter in the strict
sense; phase selection is an order of limits.  Discrete symmetries are not
covered by Coleman, Mermin--Wagner, or Hohenberg, so a two-dimensional Ising
phase and a \(1+1\)-dimensional model with a broken discrete chiral symmetry
require separate analysis.  A two-dimensional \(U(1)\) system may have a
Berezinskii--Kosterlitz--Thouless phase with power-law correlations and vortex
unbinding \cite{Berezinskii1971,KosterlitzThouless1973}, but not a nonzero
local magnetization transforming under the \(U(1)\).  Long-range exchange in a
spin model must be tested by the actual small-\(k\) behavior of
\(\sum_x J_x(1-\cos k\cdot x)\); if the second moment
\(\sum_x |x|^2|J_x|\) diverges, the quadratic denominator used above is no
longer the relevant theorem.  In the Bose theorem, gauge fields, Coulomb tails,
non-equilibrium states, boundaries, or noncompact order variables change the
field algebra or current bound and must be treated as different models.  A
large-\(N\) saddle is likewise an order-of-limits question, not a loophole in
the finite-\(N\) theorem.

Three comparison cases fix the intended reading.
\begin{enumerate}
  \item The low-temperature two-dimensional Ising model has a
  \(\mathbb Z_2\) order parameter.  Its plus/minus thermodynamic phases are
  not an exception to Mermin--Wagner--Hohenberg because the broken symmetry is
  discrete.
  \item The two-dimensional \(XY\) or compact-boson/BKT laboratory has a
  continuous \(U(1)\) symmetry, algebraic correlations in part of the phase
  diagram, and vortex physics.  It has no ordinary nonzero local \(U(1)\)
  order parameter in the thermodynamic limit.
  \item A massive relativistic \(1+1\)-dimensional \(O(N)\) sigma-model or
  Gross--Neveu-family construction may have particles in nontrivial internal
  multiplets and exact factorized scattering.  The continuous internal group
  organizes the spectrum; it is not spontaneously broken unless a separate
  local order parameter is supplied outside Coleman's hypotheses.
\end{enumerate}

\section{The \texorpdfstring{\(U(1)\)}{U(1)} Goldstone Model}

The spectral conclusion is realized explicitly by a complex scalar field with
\[
  \mathcal L
  =
  -\partial_\mu\phi^*\partial^\mu\phi
  -V(|\phi|).
\]
Figure~\ref{fig:u1-goldstone-vacuum-circle} displays the vacuum circle and the
radial/angular coordinates used to separate the massive radial excitation from
the Goldstone mode.
\begin{figure}[H]
  \centering
  \begin{tikzpicture}[>=Stealth,x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \draw[->,gray!75] (0,0) -- (0,2.25) node[above] {\(V\)};
      \draw[->,gray!75] (0,0) -- (1.95,-0.55) node[right] {\(\operatorname{Im}\phi\)};
      \draw[->,gray!75] (0,0) -- (-1.6,-0.85) node[left] {\(\operatorname{Re}\phi\)};
      \draw[thick,blue!70,domain=-1.45:1.45,smooth,variable=\x]
        plot ({\x},{0.42+0.64*(\x*\x-0.95)^2});
      \draw[dashed,blue!60] (-1.05,-0.12) arc (200:340:1.15 and 0.34);
      \draw[dashed,blue!60] (-1.05,-0.12) arc (160:20:1.15 and 0.34);
      \fill[blue!70] (0.86,-0.02) circle (1.4pt);
      \node[blue!70,align=center] at (0,-0.75)
        {\(\langle\phi\rangle=v\ee^{\ii\theta_0}\)};
      \node at (0,2.82) {exact \(U(1)\)};
    \end{scope}
    \begin{scope}[xshift=5.2cm]
      \draw[->] (-1.9,0) -- (1.9,0) node[right] {\(\operatorname{Re}\phi\)};
      \draw[->] (0,-0.15) -- (0,2.2) node[above] {\(V+\Delta V\)};
      \draw[thick,blue!70,domain=-1.55:1.55,smooth,variable=\x]
        plot ({\x},{0.16*(\x*\x-1.05)^2+0.34-0.22*\x});
      \draw[dashed,blue!45,domain=-1.35:1.35,smooth,variable=\x]
        plot ({\x},{0.16*(\x*\x-1.05)^2+0.24});
      \fill[red!70] (1.08,0.10) circle (1.6pt);
      \node[red!70,below right] at (1.08,0.10) {\(\langle\phi\rangle>0\)};
      \node at (0,2.82) {small explicit breaking};
    \end{scope}
  \end{tikzpicture}
  \caption{For a \(U(1)\)-invariant potential minimized at \(|\phi|=v\), the
  minima form a circle of vacua.  A small deformation
  \(\Delta V=-\epsilon\,\operatorname{Re}\phi\) selects one point on that
  circle.}
  \label{fig:u1-goldstone-vacuum-circle}
\end{figure}
The \(U(1)\) current is
\[
  j^\mu
  =
  -\ii(\phi^*\partial^\mu\phi-\phi\partial^\mu\phi^*).
\]
Write
\[
  \phi(x)=r(x)\ee^{\ii\theta(x)}.
\]
Then
\[
  \mathcal L
  =
  -\partial_\mu r\,\partial^\mu r
  -V(r)
  -r^2\partial_\mu\theta\,\partial^\mu\theta,
  \qquad
  j^\mu=2r^2\partial^\mu\theta .
\]
The charge \(Q=\int\dd^3x\,j^0\) generates the shift of the angular field,
\[
  [Q,\theta(x)]=\ii ,
\]
with the sign fixed by the current convention.  If \(V(r)\) is minimized at
\(r=v>0\), write \(r=v+\rho\).  The quadratic Lagrangian contains
\[
  -\partial_\mu\rho\,\partial^\mu\rho
  -\frac12V''(v)\rho^2
  -v^2\partial_\mu\theta\,\partial^\mu\theta .
\]
The radial field \(\rho\) is massive when \(V''(v)>0\), while \(\theta\) is a
massless field whose constant shifts move along the vacuum circle.

A small explicit symmetry-breaking deformation selects a phase.  For example,
\[
  \Delta V(\phi)=-\epsilon\,\operatorname{Re}\phi,
  \qquad
  \epsilon>0,
\]
chooses the vacuum with \(\theta=0\).  In the angular effective theory,
\[
  \Delta\mathcal L
  =
  \epsilon v\cos\theta
  =
  \epsilon v-\frac{\epsilon v}{2}\theta^2+O(\theta^4),
\]
so the Goldstone boson becomes a pseudo-Goldstone boson with
\[
  m_\theta^2\propto \frac{\epsilon}{v}.
\]
This is the prototype for the pion mass generated by small quark masses.

\begin{remark}[Masses in the \(U(1)\) Goldstone model]
\label{rem:u1-goldstone-model-masses}
With the mostly-plus kinetic convention used in this volume, define canonical
small-fluctuation fields near the vacuum by
\[
  h=\sqrt2\,\rho,
  \qquad
  \pi=\sqrt2\,v\,\theta .
\]
If \(V'(v)=0\) and \(V''(v)>0\), then the radial excitation has
\[
  m_h^2=\frac12V''(v),
\]
while the angular excitation is massless when the \(U(1)\) symmetry is exact.
For the explicit breaking
\(\Delta V=-\epsilon\operatorname{Re}\phi\), the angular excitation at
\(\theta=0\) has, to first order in \(\epsilon\),
\[
  m_\pi^2=\frac{\epsilon}{2v}+O(\epsilon^2).
\]
The quadratic kinetic terms are
\[
  -\partial_\mu\rho\,\partial^\mu\rho
  -v^2\partial_\mu\theta\,\partial^\mu\theta
  =
  -\frac12\partial_\mu h\,\partial^\mu h
  -\frac12\partial_\mu\pi\,\partial^\mu\pi .
\]
Expanding the symmetric potential gives
\[
  -V(v+\rho)
  =
  -V(v)-\frac12V''(v)\rho^2+O(\rho^3)
  =
  -V(v)-\frac14V''(v)h^2+O(h^3),
\]
which is the canonical mass term
\(-\frac12m_h^2h^2\) with \(m_h^2=V''(v)/2\).  No term depending on a constant
\(\theta\) is allowed by the exact shift symmetry, so \(\pi\) has no mass
term.  The explicit breaking contributes
\[
  \epsilon v\cos\theta
  =
  \epsilon v-\frac{\epsilon v}{2}\theta^2+O(\theta^4)
  =
  \epsilon v-\frac{\epsilon}{4v}\pi^2+O(\pi^4),
\]
and therefore \(m_\pi^2=\epsilon/(2v)\) in canonical normalization.
\end{remark}

\section{Chiral Symmetry of Massless QCD}

Let \(q_L^I=P_Lq^I\) and \(q_R^I=P_Rq^I\), \(I=1,\ldots,N_f\), be massless
quark fields in the fundamental representation of the color gauge group.  The
connected continuous flavor symmetry acting faithfully on local gauge-invariant
operators has Lie algebra
\[
  \mathfrak{su}(N_f)_L\oplus
  \mathfrak{su}(N_f)_R\oplus
  \mathfrak u(1)_B .
\]
For local current algebra computations this Lie algebra is represented by
\[
  SU(N_f)_L\times SU(N_f)_R\times U(1)_B .
\]
The precise global form includes finite quotients by common center actions;
those quotients matter for global background fields and are developed later.
The anomaly derived in the preceding chapter removes the axial \(U(1)_A\)
rotation from the exact continuous quantum symmetry group.

The left and right transformations act by
\[
  q_L\longmapsto Lq_L,
  \qquad
  q_R\longmapsto Rq_R,
  \qquad
  (L,R)\in SU(N_f)_L\times SU(N_f)_R .
\]

\begin{conjecture}[Chiral symmetry breaking in QCD]
\label{conj:qcd-chiral-symmetry-breaking}
For four-dimensional \(SU(N_c)\) QCD with \(N_f\) light fundamental quarks in
the range where the infrared theory is confining rather than conformal, the
infinite-volume vacuum selected by equal positive quark masses has a nonzero
chiral order parameter in the subsequent massless limit:
\begin{equation}
  \lim_{m\downarrow0}\lim_{V\to\infty}
  \langle \bar q_{R,J}q_{L,I}\rangle_{m,V}
  =
  \Sigma_0\,\delta_{IJ},
  \qquad
  \Sigma_0\neq0 .
  \label{eq:qcd-chiral-condensate-conjecture}
\end{equation}
The unbroken continuous nonabelian flavor symmetry in this phase is the
diagonal vector subgroup \(SU(N_f)_V\).
\end{conjecture}

Conjecture~\ref{conj:qcd-chiral-symmetry-breaking} is a dynamical assertion
about the QCD vacuum.  It is not presently a theorem derived from the
four-dimensional continuum QCD Lagrangian or from a completed continuum limit
of lattice QCD.  Lattice simulations and phenomenology give strong evidence
for the small-\(N_f\) theory, while semiclassical instanton calculations do
not supply a controlled derivation of the infrared condensate.  The pion
effective theory below is the low-energy theory conditional on this chiral
breaking phase.

In the phase described by
Conjecture~\ref{conj:qcd-chiral-symmetry-breaking}, the bilinear condensate has
an orientation that can be chosen as
\[
  \Sigma_{IJ}
  =
  \langle \bar q_{R,J}q_{L,I}\rangle
  =
  \Sigma_0\,\delta_{IJ},
\]
so that the stabilizer is the diagonal subgroup.  More explicitly, before
choosing the orientation of the vacuum, write
\[
  A_{IJ}:=\langle\bar q_{R,J}q_{L,I}\rangle .
\]
Under \(SU(N_f)_L\times SU(N_f)_R\),
\[
  A\longmapsto R^\dagger A L .
\]
A choice of \(L\) and \(R\) can diagonalize \(A\).  If the vacuum is invariant
under the non-chiral subgroup, then \(A\) is proportional to the identity.
The unbroken transformations are precisely \(L=R\).  The breaking pattern is
therefore
\begin{equation}
  SU(N_f)_L\times SU(N_f)_R
  \longrightarrow
  SU(N_f)_V .
  \label{eq:qcd-chiral-breaking-pattern}
\end{equation}
Given this dynamical input, the number of Goldstone bosons is \(N_f^2-1\).

\begin{definition}[QCD Goldstone target]
\label{def:qcd-chiral-goldstone-datum}
Conditional on the breaking pattern
\eqref{eq:qcd-chiral-breaking-pattern}, the Goldstone target is
\[
  \frac{SU(N_f)_L\times SU(N_f)_R}{SU(N_f)_V}
  \simeq SU(N_f).
\]
The coordinate \(U\in SU(N_f)\) is chosen with left flavor as the row index
and right flavor as the column index.  With this index convention it
parameterizes the condensate orientation relative to the reference condensate
\(U=1\), and the left and right flavor groups act by
\[
  U\longmapsto LUR^{-1}.
\]
This is the index convention used by the mass spurion and the pion effective
action below; it is the same physical orientation as the bilinear matrix
\(A_{IJ}=\langle\bar q_{R,J}q_{L,I}\rangle\) above, written with the opposite
choice of displayed row and column indices.
The diagonal subgroup \(L=R\) fixes the reference orientation \(U=1\) and acts
on the tangent space at \(U=1\) by the adjoint representation of
\(\mathfrak{su}(N_f)\).
\end{definition}

At \(\theta=0\), the non-chiral flavor symmetry is constrained by the
positivity structure of the Euclidean vectorlike theory.  The statement has a
precise theorem boundary.

\begin{hypothesis}[Euclidean measure needed for Vafa--Witten positivity]
\label{hyp:vafa-witten-measure-data}
For the vectorlike theory with equal positive quark masses \(m>0\), assume that
there is a family of gauge-invariant Euclidean finite-regulator measures
\(\dd\nu_{m,\Lambda}(A)\).  If a continuum statement is claimed, assume in
addition that the gauge-invariant Schwinger functions and the source
functionals for the order parameters used below have a continuum limit, denoted
\(\dd\nu_m(A)\) in this paragraph, with the following properties:
\begin{enumerate}[label=\roman*.]
\item expectations of gauge-invariant polynomially bounded fermion bilinears and
their source deformations are defined at the regulator level and converge in
the limit in which the Vafa--Witten conclusion is asserted;
\item the full regulator measure, including the gauge action and the vectorlike
fermion discretization or determinant, is reflection positive for a Euclidean
time reflection, and this positivity is preserved by the limiting Schwinger
functions.  Thus OS reconstruction applies to the observables used in the
argument;
\item the Dirac operator and fermion regulator have vectorlike determinant
positivity in each fixed gauge-field background.  In the continuum notation
this is the anti-Hermiticity and eigenvalue pairing statement
\[
  \det(\not D_A+m)
  =
  \prod_{\lambda>0}(\lambda^2+m^2)\,m^{n_0(A)}
  \geq0;
\]
on a lattice it is replaced by the corresponding positivity property of the
chosen vectorlike fermion action;
\item the infinite-volume and small-mass limits used to select a phase exist
for the order parameters under consideration.
\end{enumerate}
\end{hypothesis}

At finite lattice spacing, these hypotheses are properties to verify for the
chosen lattice gauge and fermion action.  The Wilson plaquette gauge action is
reflection positive in the sense described in
Chapter~\ref{ch:lattice-yang-mills-nonperturbative-definition}, and standard
vectorlike fermion discretizations must be checked with their own
reflection-positivity and determinant-positivity conditions.  These properties
are not automatic for an arbitrary regulator or gauge-fixed expression.  For
the four-dimensional continuum gauge theory, the existence of \(\dd\nu_m(A)\)
with the required locality, reflection positivity, and scaling limits is part
of the Yang--Mills/QCD construction problem; the finite-lattice positivity
argument is therefore not, by itself, a construction of continuum QCD.

Under Hypothesis~\ref{hyp:vafa-witten-measure-data}, the determinant factor in
each fixed gauge-field background is nonnegative, and for equal masses the
fermion determinant and gauge measure are invariant under \(SU(N_f)_V\).  A
vector-flavor-breaking external source may then be averaged over the compact
vector flavor group without decreasing the Euclidean functional integral.  In
the infinite-volume limit this convexity/positivity comparison rules out a
lower vacuum energy for a vector-flavor-breaking orientation.  Taking the
small-mass limit therefore selects, when the limit exists, a vacuum preserving
\(SU(N_f)_V\).  This is the Vafa--Witten positivity argument in the form
needed here.  It does not by itself prove the existence of the chiral
condensate; it constrains the orientation of the condensate once the
vectorlike massive regulator and its reflection-positive measure have been
constructed.

The complex phase of \(\Sigma_0\) is tied to the topological angle.  A
flavor-singlet axial rotation would rotate \(\Sigma_0\), but the same
transformation is anomalous and shifts \(\theta\).  Thus the invariant structure
are the condensate orientation together with the anomaly-invariant
combination of \(\theta\) and the quark-mass phase discussed in the preceding
chapter.

\section{Topological Susceptibility and the Singlet Axial Direction}
\label{sec:qcd-topological-susceptibility-singlet}

The nonabelian pion field below takes values in \(SU(N_f)\), because the
axial \(U(1)_A\) rotation is anomalous at finite \(N_c\).  At large \(N_c\)
the anomaly is suppressed in the singlet meson sector in a precise sense:
the pure Yang--Mills topological susceptibility remains \(O(1)\), while
\(f_\pi^2=O(N_c)\).  The resulting singlet mass is therefore \(O(N_c^{-1})\).
This paragraph is a conditional large-\(N_c\) matching statement, not a proof
of the pure Yang--Mills continuum limit or of the full QCD chiral phase.

\begin{hypothesis}[Large-\(N_c\) singlet theta matching]
\label{hyp:large-n-singlet-theta-datum}
Fix \(N_f\) while \(N_c\to\infty\), and assume the following inputs.
\begin{enumerate}
\item Pure \(SU(N_c)\) Yang--Mills has a topological susceptibility
  \(\chi_{\rm YM}\) in the sense of
  Definition~\ref{def:topological-susceptibility}, with
  \(\chi_{\rm YM}=O(N_c^0)\).
\item In the branch near \(\vartheta=0\), the pure Yang--Mills vacuum energy
  has the local expansion
  \begin{equation}
    E_{\rm YM}(\vartheta)
    =
    E_{\rm YM}(0)
    +\frac12\chi_{\rm YM}\vartheta^2
    +O(\vartheta^4/N_c^2),
    \label{eq:large-n-ym-theta-local-expansion}
  \end{equation}
  while \(2\pi\)-periodicity is restored by the usual branch label
  \(\vartheta\mapsto \vartheta+2\pi k\).
\item The chiral effective field may be enlarged locally from \(SU(N_f)\) to
  \(U(N_f)\) near the identity:
  \begin{equation}
    \widehat U
    =
    \exp\!\left(\frac{2\ii}{f_\pi}
      (\eta_0T^0+\pi^aT^a)\right),
    \qquad
    T^0=\frac{\mathbf 1_{N_f}}{\sqrt{2N_f}},
    \qquad
    \operatorname{Tr}(T^AT^B)=\frac12\delta^{AB}.
    \label{eq:singlet-u-nf-field}
  \end{equation}
  The leading kinetic term is
  \begin{equation}
    \frac{f_\pi^2}{4}
    \operatorname{Tr}(\partial_\mu\widehat U\partial^\mu\widehat U^\dagger),
    \label{eq:singlet-u-nf-kinetic-term}
  \end{equation}
  so that \(\eta_0\) is canonically normalized.
\item To leading order in the fixed-\(N_f\) large-\(N_c\) expansion and at
  vanishing quark masses, the theta dependence enters the singlet effective
  potential through
  \begin{equation}
    \vartheta
    =
    \theta-\ii\log\det\widehat U
    \label{eq:singlet-theta-combination}
  \end{equation}
  on the branch near \(\widehat U=1\).  With the convention
  \(U\mapsto LUR^{-1}\), a common axial change of variables sends
  \(\theta\mapsto\theta-2N_f\alpha\) and
  \(\ii\log\det\widehat U\mapsto
  \ii\log\det\widehat U-2N_f\alpha\), so this is the invariant coordinate.
  When the mass term aligns \(\widehat U\) with the phase of \(M\), it reduces
  to the microscopic strong-CP phase \(\theta+\arg\det M\).
\end{enumerate}
\end{hypothesis}

\paragraph{Witten--Veneziano mass formula from the large-\(N_c\) matching input.}
Under Hypothesis~\ref{hyp:large-n-singlet-theta-datum}, the singlet field
\(\eta_0\) has, in the chiral limit and at leading fixed-\(N_f\) large
\(N_c\),
\begin{equation}
  m_{\eta_0}^2
  =
  \frac{2N_f}{f_\pi^2}\chi_{\rm YM}
  +O(N_c^{-2}) .
  \label{eq:witten-veneziano-chiral-limit}
\end{equation}
Equivalently, because \(f_\pi^2=O(N_c)\) while
\(\chi_{\rm YM}=O(1)\), the singlet mass squared is \(O(N_c^{-1})\).

Since the non-singlet generators are traceless,
\begin{equation}
  -\ii\log\det\widehat U
  =
  \frac{2\eta_0}{f_\pi}\operatorname{Tr}T^0
  =
  \frac{\sqrt{2N_f}}{f_\pi}\eta_0
  \label{eq:singlet-determinant-phase}
\end{equation}
on the branch near the identity.  Substituting
\eqref{eq:singlet-theta-combination} into the local Yang--Mills branch
\eqref{eq:large-n-ym-theta-local-expansion} gives the singlet potential
\[
  V_{\rm singlet}(\eta_0;\theta)
  =
  \frac12\chi_{\rm YM}
  \left(
    \theta+\frac{\sqrt{2N_f}}{f_\pi}\eta_0
  \right)^2
  +O(N_c^{-2})
  \]
in the effective Lagrangian.  At fixed \(\theta=0\), the quadratic term is
\[
  \frac12
  \left(\frac{2N_f\chi_{\rm YM}}{f_\pi^2}\right)\eta_0^2 .
  \]
The kinetic term
\eqref{eq:singlet-u-nf-kinetic-term} gives
\(\frac12\partial_\mu\eta_0\partial^\mu\eta_0\), because
\(\operatorname{Tr}(T^0T^0)=1/2\).  The displayed coefficient is therefore the
mass squared.

\paragraph{Theta-curvature matrix and the screening direction.}
The finite-regulator cumulant identity
\eqref{eq:finite-volume-topological-susceptibility} fixes the sign of the
pure Yang--Mills curvature:
\[
  \chi_{\rm YM}=E_{\rm YM}''(0)\ge0
\]
when the limit exists in a CP-invariant branch.  In the local singlet chart set
\[
  a:=\frac{\sqrt{2N_f}}{f_\pi},
  \qquad
  V(\theta,\eta_0)
  =
  \frac12\chi_{\rm YM}(\theta+a\eta_0)^2 .
\]
The Hessian in the coordinate pair \((\theta,\eta_0)\) is the rank-one matrix
\begin{equation}
  H_{\theta,\eta}
  =
  \chi_{\rm YM}
  \begin{pmatrix}
    1&a\\
    a&a^2
  \end{pmatrix}.
  \label{eq:theta-eta0-curvature-matrix}
\end{equation}
Thus \(H_{\theta,\eta}(-a,1)^T=0\): shifting \(\theta\) and \(\eta_0\) so that
\(\theta+a\eta_0\) is fixed is the massless-quark screening direction.
By contrast, the fixed-\(\theta\) curvature in the \(\eta_0\) direction is
\[
  \partial_{\eta_0}^2V|_{\theta=0}
  =
  a^2\chi_{\rm YM}
  =
  \frac{2N_f}{f_\pi^2}\chi_{\rm YM},
\]
which is exactly the mass coefficient in
\eqref{eq:witten-veneziano-chiral-limit}.  The mixed derivative
\(\partial_\theta\partial_{\eta_0}V=a\chi_{\rm YM}\) records the sign of the
anomaly-invariant coordinate
\(\theta-\ii\log\det\widehat U\); reversing that sign would give the same
positive mass but the opposite singlet screening convention.

\begin{controlledapproximation}[Residual-controlled Witten--Veneziano matching]
\label{ca:residual-controlled-witten-veneziano}
The preceding rank-one matrix is the leading local Hessian in the chiral,
fixed-\(N_f\), large-\(N_c\) matching window.  If the comparison is made in
the exact massless full-QCD theory, with a continuum and regulator convention
which preserves the anomalous axial Ward identity, the first residual cannot
be an arbitrary symmetric \(2\times2\) matrix.  With at least one exactly
massless quark the \(\theta\)-angle is removable by an anomalous chiral
rotation, so the full vacuum energy is independent of \(\theta\).  Therefore
the local residual must preserve the screening vector:
\[
  H_{\theta,\eta}^{\rm loc}(-a,1)^T=0 .
\]
In the two-variable local Hessian this condition forces the residual to stay
in the same anomaly-invariant coordinate.  Thus write
\begin{equation}
  H_{\theta,\eta}^{\rm loc}
  =
    \chi_{\rm eff}
    \begin{pmatrix}1&a\\ a&a^2\end{pmatrix}
  =
    (\chi_{\rm YM}+\delta\chi)
    \begin{pmatrix}1&a\\ a&a^2\end{pmatrix},
  \qquad
  a=\frac{\sqrt{2N_f}}{f_\pi},
  \label{eq:wv-residual-local-hessian}
\end{equation}
with \(\chi_{\rm eff}>0\) in the selected local branch.  The scalar
\(\delta\chi\) records the declared large-\(N_c\), branch-selection, and local
counterterm uncertainty in the same renormalization convention as
\(\chi_{\rm YM}\); it shifts the fixed-\(\theta\) singlet curvature and the
mixed curvature together.  At fixed \(\theta\), the singlet potential
curvature in the leading field coordinate is
\begin{equation}
  H_{\eta\eta}^{\rm loc}
  =
  a^2\chi_{\rm eff}
  =
  \frac{2N_f}{f_\pi^2}(\chi_{\rm YM}+\delta\chi).
  \label{eq:wv-residual-singlet-curvature}
\end{equation}
But minimizing the same local quadratic form over the singlet field gives
\begin{equation}
  \chi_{\rm QCD}^{m=0,{\rm loc}}
  =
  \chi_{\rm eff}
  -
  \frac{(a\chi_{\rm eff})^2}{a^2\chi_{\rm eff}}
  =
  0 .
  \label{eq:wv-residual-schur-curvature}
\end{equation}
The exact zero in \eqref{eq:wv-residual-schur-curvature} is part of the
massless matching condition, not an optional cancellation imposed after the
fact.

A physical pole mass also requires the kinetic normalization of the singlet
field.  At the same order at which \(\delta\chi\) is kept, write the local
quadratic kinetic term in the \(\eta_0\) coordinate as
\begin{equation}
  \frac12 Z_0\,\partial_\mu\eta_0\partial^\mu\eta_0,
  \qquad
  F_0^2:=Z_0f_\pi^2 .
  \label{eq:wv-singlet-kinetic-residual}
\end{equation}
Then the mass statement is the one-field generalized eigenvalue
\begin{equation}
  m_{\eta_0,{\rm phys}}^2
  =
  \frac{H_{\eta\eta}^{\rm loc}}{Z_0}
  =
  \frac{2N_f}{F_0^2}(\chi_{\rm YM}+\delta\chi),
  \label{eq:wv-residual-physical-singlet-mass}
\end{equation}
while \eqref{eq:wv-residual-singlet-curvature} is the corresponding potential
curvature in the leading field coordinate.  Thus the usable uncertainty
statement is: if \(|\delta\chi|\le r_\chi\), \(|Z_0-1|\le r_Z<1\), and the
leading comparison mass is \(m_{\rm WV,0}^2=2N_f\chi_{\rm YM}/f_\pi^2\), then
\begin{equation}
  \left|
    m_{\eta_0,{\rm phys}}^2-m_{\rm WV,0}^2
  \right|
  \le
  \frac{2N_f}{f_\pi^2}\,
  \frac{r_\chi+\chi_{\rm YM}r_Z}{1-r_Z},
  \qquad
  \chi_{\rm QCD}^{m=0,{\rm loc}}\equiv0 .
  \label{eq:wv-residual-mass-bound}
\end{equation}
The first term in the numerator is the potential-curvature uncertainty; the
second is the independent singlet decay-constant or wavefunction uncertainty.
The anomalous Ward identity fixes the screening coordinate of the potential.  It
does not fix \(Z_0\).

If one intentionally leaves this massless Ward-compatible setting, for example
by keeping finite quark masses or by introducing an explicit symmetry-breaking
regulator/counterterm error, one may add a separate Ward-breaking matrix
\(B=(b_{ij})\):
\begin{equation}
  H_{\theta,\eta}^{\rm break}
  =
  \chi_{\rm eff}
  \begin{pmatrix}1&a\\a&a^2\end{pmatrix}
  +
  \begin{pmatrix}
    b_{\theta\theta}&b_{\theta\eta}\\
    b_{\theta\eta}&b_{\eta\eta}
  \end{pmatrix}.
  \label{eq:wv-ward-breaking-hessian}
\end{equation}
For \(a^2\chi_{\rm eff}+b_{\eta\eta}>0\), the Schur complement of
\eqref{eq:wv-ward-breaking-hessian} is
\begin{equation}
  \chi_{\rm break}^{\rm loc}
  =
  \frac{
    \chi_{\rm eff}b_{\eta\eta}
    +a^2\chi_{\rm eff}b_{\theta\theta}
    +b_{\theta\theta}b_{\eta\eta}
    -2a\chi_{\rm eff}b_{\theta\eta}
    -b_{\theta\eta}^2}
  {a^2\chi_{\rm eff}+b_{\eta\eta}} .
  \label{eq:wv-ward-breaking-schur-diagnostic}
\end{equation}
This last quantity is a diagnostic for the declared breaking source or for
finite-mass physics.  It is not the theta curvature of the exactly massless,
symmetry-preserving full-QCD theory.  Equations
\eqref{eq:wv-residual-local-hessian}--\eqref{eq:wv-residual-mass-bound} are the
local bookkeeping needed before comparing a measured \(\eta'\) mass, a lattice
\(\chi_{\rm YM}\), and a full-QCD topological susceptibility.  They do not
prove the Yang--Mills continuum limit, the large-\(N_c\) expansion, or the
absence of branch crossings; they state how those inputs enter the observable
singlet curvature once a local branch and counterterm convention have been
fixed.

With nonzero quark masses the same caveat becomes a matrix statement.  For the
neutral fields retained in the effective theory, write
\[
  \mathcal L_2
  =
  \frac12 K_{AB}\partial_\mu\varphi^A\partial^\mu\varphi^B
  -\frac12 H_{AB}\varphi^A\varphi^B+\cdots .
\]
The physical pseudoscalar masses are the generalized eigenvalues
\begin{equation}
  \det(H-m^2K)=0 ,
  \label{eq:wv-neutral-generalized-eigenvalue}
\end{equation}
so \(\eta_0\)-octet potential mixing, kinetic mixing, and the singlet
wavefunction residual must be separated from the pure-Yang--Mills theta
curvature.  The scalar formula
\eqref{eq:wv-residual-physical-singlet-mass} is the chiral, unmixed,
single-field specialization of \eqref{eq:wv-neutral-generalized-eigenvalue}.
\end{controlledapproximation}

\begin{remark}[Why this does not contradict massless-quark theta independence]
\label{rem:massless-quark-theta-independence-wv}
If the quark masses vanish and \(\eta_0\) is included as a dynamical field,
the vacuum minimizes
\(
\theta+\sqrt{2N_f}\eta_0/f_\pi
\).
Thus the full massless QCD vacuum energy is locally independent of
\(\theta\), and its topological susceptibility vanishes, while the curvature
in the singlet field direction at fixed \(\theta\) is nonzero.  The
susceptibility in \eqref{eq:witten-veneziano-chiral-limit} is the pure
Yang--Mills susceptibility \(\chi_{\rm YM}\), not the full massless-QCD
susceptibility.  This is the local Hessian form of the anomalous axial Ward
identity and of the Leutwyler--Smilga small-mass constraint
\cite{LeutwylerSmilga1992}: for positive small quark masses,
\(\chi_{\rm QCD}=\Sigma(\sum_i m_i^{-1})^{-1}+O(m^2)\), so the curvature
vanishes when any one quark mass is taken to zero.
\end{remark}

\begin{controlledapproximation}[Instanton determinant as a chiral spurion
potential]
\label{ca:instanton-determinant-chiral-spurion-potential}
Assume, in addition to the chiral phase hypothesis above, that the regulated
one-instanton amplitude of
Chapter~\ref{ch:volume-ii-20} has been matched at a scale \(\mu\)
to a local color-singlet determinant source and that the instanton and
anti-instanton sectors form a dilute finite-activity expansion in the sense of
\eqref{ca:dilute-instanton-gas-amplitude}.  Denote the resulting
renormalized activity in the chosen chiral scheme by
\(\zeta_\chi(\mu)>0\).  Then the zero-mode determinant and its
anti-instanton conjugate induce the local singlet potential
\begin{equation}
  \begin{aligned}
    V_{\rm inst}^{\chi}(\widehat U;\theta)
    &=
    2\zeta_\chi
    \left[
      1-\cos\!\left(\theta-\ii\log\det\widehat U\right)
    \right] \\
    &\quad
    +\hbox{higher-derivative and higher-mass terms}
    +\hbox{multi-instanton terms},
  \end{aligned}
  \label{eq:instanton-chiral-determinant-potential}
\end{equation}
on the same local branch as
\eqref{eq:singlet-theta-combination}.  Equivalently, in the
\(\widehat U\) convention of \eqref{eq:singlet-u-nf-field}, the corresponding
Lagrangian vertex is proportional to
\[
  \zeta_\chi\,
  \ee^{\ii\theta}\det\widehat U
  +
  \zeta_\chi\,
  \ee^{-\ii\theta}\det\widehat U^\dagger,
\]
up to a field-independent constant in the potential.  If the microscopic
zero-mode determinant is written with the conjugate row/column convention,
the displayed vertex is obtained by the corresponding transpose/conjugation
conversion; the sign used here is fixed by the singlet coordinate
\eqref{eq:singlet-theta-combination}.

In the \(\eta_0\) chart and sign convention of
\eqref{eq:theta-eta0-curvature-matrix}, set
\[
  a=\frac{\sqrt{2N_f}}{f_\pi},
  \qquad
  \theta-\ii\log\det\widehat U
  =
  \theta+a\eta_0
  \quad\hbox{on the chosen exponential branch}.
\]
Expanding \eqref{eq:instanton-chiral-determinant-potential} gives
\begin{equation}
  V_{\rm inst}^{\chi}(\theta,\eta_0)
  =
  \zeta_\chi(\theta+a\eta_0)^2
  +O((\theta+a\eta_0)^4),
  \label{eq:instanton-chiral-determinant-quadratic}
\end{equation}
so its local curvature matrix is
\begin{equation}
  H_{\theta,\eta}^{\rm inst}
  =
  2\zeta_\chi
  \begin{pmatrix}
    1&a\\
    a&a^2
  \end{pmatrix}.
  \label{eq:instanton-chiral-curvature-matrix}
\end{equation}
The same coefficient \(2\zeta_\chi\) is the topological susceptibility of the
conditional dilute gas in
\eqref{eq:dilute-instanton-gas-cumulants}.  Therefore a regime in which
\(\zeta_\chi\) is actually calculable would give the singlet mass contribution
\begin{equation}
  \Delta m_{\eta_0,{\rm inst}}^2
  =
  \frac{2N_f}{f_\pi^2}\,(2\zeta_\chi)
  =
  \frac{4N_f\zeta_\chi}{f_\pi^2}.
  \label{eq:instanton-dilute-singlet-mass-contribution}
\end{equation}
This is the dilute-instanton analogue of
\eqref{eq:witten-veneziano-chiral-limit}, with
\(\chi_{\rm YM}\) replaced by the finite activity susceptibility
\(\chi_{\rm top}^{\rm dig}=2\zeta_\chi\).

The qualification is part of the statement.  In zero-temperature confining
QCD the large-\(\rho\) part of the one-instanton activity and the
multi-instanton interaction expansion are not controlled by the BPST
calculation alone.  Thus \eqref{eq:instanton-chiral-determinant-potential} is
a valid chiral coordinate for a declared finite-activity semiclassical window,
such as a weakly coupled Higgsed or sufficiently screened regime, but it is
not a derivation of the physical \(\eta'\) mass from isolated instantons in
ordinary QCD.  The large-\(N_c\) Witten--Veneziano input above instead uses
the pure Yang--Mills susceptibility as a nonperturbative datum.
\end{controlledapproximation}

\begin{controlledapproximation}[Instanton-to-Witten--Veneziano comparison
window]
\label{ca:instanton-wv-comparison-window}
The similarity between \eqref{eq:instanton-chiral-curvature-matrix} and
\eqref{eq:theta-eta0-curvature-matrix} is useful only after the two
curvatures have been placed in the same scheme and branch.  Set
\[
  \chi_{\rm inst}:=2\zeta_\chi,
  \qquad
  a=\frac{\sqrt{2N_f}}{f_\pi}.
\]
The dilute determinant gives the conditional finite-activity curvature
\(\chi_{\rm inst}\) of its selected instanton branch.  The
Witten--Veneziano formula uses the pure Yang--Mills curvature
\(\chi_{\rm YM}\).  A comparison window therefore requires, as an extra
physical input, a same-scheme curvature estimate
\begin{equation}
  |\chi_{\rm YM}-\chi_{\rm inst}|\le \epsilon_\chi
  \label{eq:instanton-wv-curvature-distance}
\end{equation}
together with the kinetic and mixing residuals needed to turn a local
curvature into a pole mass.  With singlet kinetic coefficient \(Z_0\), the
chiral, unmixed comparison masses are
\[
  m_{0,{\rm WV}}^2
  =
  \frac{2N_f\chi_{\rm YM}}{Z_0f_\pi^2},
  \qquad
  m_{0,{\rm inst}}^2
  =
  \frac{2N_f\chi_{\rm inst}}{Z_0f_\pi^2},
\]
and \eqref{eq:instanton-wv-curvature-distance} gives
\begin{equation}
  |m_{0,{\rm WV}}^2-m_{0,{\rm inst}}^2|
  \le
  \frac{2N_f}{Z_0f_\pi^2}\epsilon_\chi
  +\epsilon_{\rm br}+\epsilon_{\rm mix}.
  \label{eq:instanton-wv-mass-comparison-bound}
\end{equation}
Here \(\epsilon_{\rm br}\) denotes branch, large-\(N_c\), and local
counterterm mismatch not already included in \(\epsilon_\chi\), while
\(\epsilon_{\rm mix}\) denotes the octet-mixing and higher-order chiral
remainder in the chosen observable channel.

The full-QCD massless susceptibility is a different operation.  For either
\(\chi=\chi_{\rm YM}\) or \(\chi=\chi_{\rm inst}\), the local Hessian
\[
  H(\chi)
  =
  \chi
  \begin{pmatrix}
    1&a\\
    a&a^2
  \end{pmatrix}
\]
has the Schur complement
\begin{equation}
  H_{\theta\theta}
  -
  H_{\theta\eta}H_{\eta\eta}^{-1}H_{\eta\theta}
  =
  \chi-\frac{(a\chi)^2}{a^2\chi}
  =
  0 .
  \label{eq:instanton-wv-massless-schur-zero}
\end{equation}
Thus the instanton branch curvature, like the Witten--Veneziano curvature, is
a fixed-singlet or conditional-branch curvature before the dynamical singlet
field is minimized.  Substituting the full massless-QCD susceptibility
\(\chi_{\rm QCD}=0\) into the Witten--Veneziano mass formula would erase the
fixed-\(\theta\) singlet curvature and is the wrong observable comparison.
Conversely, substituting \(2\zeta_\chi\) for \(\chi_{\rm YM}\) in the physical
\(\eta'\) trace relation is justified only after
\eqref{eq:instanton-wv-curvature-distance} and the pole-mass residuals have
been supplied.

For example, in the three-flavor trace observable below, the instanton branch
may be used as a tested approximant only in the form
\begin{equation}
  \left|
    m_{\eta'}^2+m_\eta^2-2m_K^2
    -
    \frac{2N_f}{f_\pi^2}\chi_{\rm inst}
  \right|
  \le
  \frac{2N_f}{f_\pi^2}\epsilon_\chi
  +R_{\rm chiral}+R_{\rm kin}+R_{\rm mix}.
  \label{eq:instanton-wv-trace-comparison-budget}
\end{equation}
This is the observable form of the physics distinction: the Chapter~20
instanton calculation supplies a source-saturated semiclassical amplitude and,
in finite-activity regimes, a conditional theta curvature.  The
Witten--Veneziano relation supplies a large-\(N_c\) pure-glue curvature input
and then a generalized-eigenvalue calculation in the pseudoscalar channel.
Finite-momentum \(U(1)_A\) source kernels, such as the \(\pi\)-\(\delta\)
diagnostics of the instanton section, are related probes of the same
zero-mode determinant but are not themselves topological susceptibilities or
\(\eta'\) pole masses until the corresponding spectral and LSZ matching has
been done.
\end{controlledapproximation}

For nonzero light-quark masses the singlet field mixes with the octet
pseudoscalars.  This mixing is part of the susceptibility argument at the
level of observables: it is the finite mass-matrix step which turns the pure
Yang--Mills theta curvature into a statement about the physical pseudoscalar
spectrum.
For \(N_f=3\), take the isospin limit
\[
  M=\operatorname{diag}(m,m,m_s),
  \qquad
  m_\pi^2=2B_0m,
  \qquad
  m_K^2=B_0(m+m_s),
\]
and use the local \(U(3)\) coordinates of
\eqref{eq:singlet-u-nf-field} with
\[
  T^3=\frac12\operatorname{diag}(1,-1,0),
  \qquad
  T^8=\frac1{2\sqrt3}\operatorname{diag}(1,1,-2),
  \qquad
  T^0=\frac1{\sqrt6}\mathbf 1_3 .
\]
Expanding the leading mass spurion
\(\frac{f_\pi^2}{4}\operatorname{Tr}(\chi\widehat U^\dagger+
\widehat U\chi^\dagger)\) to quadratic order gives the neutral mass matrix
\[
  (M^2_{\rm mass})_{AB}
  =
  4B_0\,\frac12
  \operatorname{Tr}\bigl(M\{T^A,T^B\}\bigr),
  \qquad A,B\in\{3,8,0\}.
\]
The isospin generator decouples:
\[
  M^2_{33}=m_\pi^2,
  \qquad
  M^2_{38}=M^2_{30}=0 .
\]
Adding the anomalous singlet curvature
\[
  m_0^2:=\frac{2N_f}{f_\pi^2}\chi_{\rm YM}
  \qquad (N_f=3)
\]
only in the \(00\) entry gives the \((\eta_8,\eta_0)\) block
\begin{equation}
  M^2_{\eta_8\eta_0}
  =
  \begin{pmatrix}
    \frac{4m_K^2-m_\pi^2}{3}
    &
    -\frac{2\sqrt{2}}{3}(m_K^2-m_\pi^2)
    \\[0.4em]
    -\frac{2\sqrt{2}}{3}(m_K^2-m_\pi^2)
    &
    \frac{2m_K^2+m_\pi^2}{3}+m_0^2
  \end{pmatrix}.
  \label{eq:eta-eta-prime-leading-mass-matrix}
\end{equation}
Changing the sign convention for \(\eta_8\) changes the sign of the off-diagonal
entry, but not the eigenvalues.  If those eigenvalues are denoted
\(m_\eta^2\) and \(m_{\eta'}^2\) at this leading order, the trace gives
\begin{equation}
  m_{\eta'}^2+m_\eta^2-2m_K^2
  =
  m_0^2
  =
  \frac{2N_f}{f_\pi^2}\chi_{\rm YM}.
  \label{eq:eta-eta-prime-trace-witten-veneziano}
\end{equation}
The determinant records the same mixing data in the product of the two
eigenvalues:
\begin{equation}
  m_\eta^2m_{\eta'}^2
  =
  m_\pi^2(2m_K^2-m_\pi^2)
  +
  \frac{m_0^2}{3}(4m_K^2-m_\pi^2).
  \label{eq:eta-eta-prime-determinant-ledger}
\end{equation}
Thus the usual phenomenological form
\[
  m_{\eta'}^2+m_\eta^2-2m_K^2
  =
  \frac{2N_f}{f_\pi^2}\chi_{\rm YM}
  +\hbox{controlled corrections}
\]
requires the quark-mass potential, the normalization of \(f_\pi\), and the
large-\(N_c\) and chiral remainders to be stated in the same scheme.  The
chiral-limit result above fixes the normalization; the finite matrix
\eqref{eq:eta-eta-prime-leading-mass-matrix} shows how that curvature enters a
physical mass spectrum once light-quark masses are restored.

Choose Hermitian flavor generators \(T^a\) with
\(\operatorname{Tr}(T^aT^b)=\frac12\delta^{ab}\).  These \(T^a\) are flavor
generators used to parameterize the chiral target \(SU(N_f)\); they are not
the color gauge generators \(t^a\) of the Yang--Mills chapters, where the
default convention is \(\operatorname{tr}(t^a t^b)=\delta^{ab}\).  The
half-trace flavor convention is kept here because, together with the factor
\(2/f_\pi\) below, it gives the canonical leading pion coordinates and is
compatible with the axial-current normalization stated below.
The pion field is the
\(SU(N_f)\)-valued matrix
\begin{equation}
  U(x)
  =
  \exp\left\{\frac{2\ii}{f_\pi}\pi^a(x)T^a\right\},
  \qquad
  U(x)\in SU(N_f),
  \label{eq:pion-matrix-field}
\end{equation}
transforming as
\begin{equation}
  U(x)
  \longmapsto
  L\,U(x)\,R^{-1}.
  \label{eq:pion-chiral-transformation}
\end{equation}
The axial-current normalization is
\[
  \bra{\vac}A^{a\mu}(0)\ket{\pi^b(p)}
  =
  \ii f_\pi p^\mu\delta^{ab}.
\]
With this convention, the leading two-derivative chiral Lagrangian is
\begin{equation}
  \mathcal L_2
  =
  -\frac{f_\pi^2}{4}
  \operatorname{Tr}
  \bigl(D_\mu U^\dagger D^\mu U\bigr),
  \label{eq:leading-chiral-lagrangian}
\end{equation}
where the covariant derivative in external left and right flavor backgrounds
\(\ell_\mu,r_\mu\) is
\begin{equation}
  D_\mu U
  =
  \partial_\mu U-\ii\ell_\mu U+\ii U r_\mu .
  \label{eq:pion-external-covariant-derivative}
\end{equation}
The minus sign in \eqref{eq:leading-chiral-lagrangian} is the mostly-plus
metric convention used throughout this volume.  Expanding
\eqref{eq:pion-matrix-field} at \(\ell_\mu=r_\mu=0\) gives
\[
  \operatorname{Tr}(\partial_\mu U^\dagger\partial^\mu U)
  =
  \frac2{f_\pi^2}\,
  \partial_\mu\pi^a\partial^\mu\pi^a
  +O(\pi^3),
\]
and hence the canonical leading kinetic term
\[
  \mathcal L_2
  =
  -\frac12\partial_\mu\pi^a\partial^\mu\pi^a+\cdots .
\]
The same expansion also fixes the normalization of the derivative
self-interactions.

\section{Soft-Pion Ward Identity and the Adler Zero}
\label{sec:soft-pion-adler-zero}

The axial-current normalization above fixes the normalization of the one-pion
pole in current correlators.  Let \(\Pi^a(x)\) be a local pion interpolating
field in the low-energy theory, normalized by
\[
  \bra{\vac}\Pi^a(0)\ket{\pi^b(p)}
  =
  Z_\pi^{1/2}\delta^{ab},
  \qquad p^2=0,
\]
and let
\[
  G_a^\mu(x;x_1,\ldots,x_n)
  =
  \bra{\vac}
  T\{A_a^\mu(x)\Pi^{a_1}(x_1)\cdots\Pi^{a_n}(x_n)\}
  \ket{\vac}
\]
be the time-ordered distribution with one non-singlet axial current.  In the
massless theory, and with the anomalous \(U(1)_A\) current excluded, the local
Ward identity is
\begin{equation}
\begin{aligned}
  \partial_\mu G_a^\mu(x;x_1,\ldots,x_n)
  &=
  -\ii\sum_{j=1}^n
  \delta^{(4)}(x-x_j)
\\
  &\quad\times
  \bra{\vac}
  T\{\Pi^{a_1}(x_1)\cdots
  \delta_A^a\Pi^{a_j}(x_j)\cdots\Pi^{a_n}(x_n)\}
  \ket{\vac}.
\end{aligned}
\label{eq:soft-pion-current-ward-identity}
\end{equation}
Here \(\delta_A^a\) denotes the infinitesimal transformation generated by the
broken axial charge \(Q_A^a\).  Equation
\eqref{eq:soft-pion-current-ward-identity} is an identity of distributions:
the right side consists of contact terms because the divergence of a conserved
current inside a time-ordered product differentiates the step functions.

Fourier transform in \(x\) with momentum \(q\) entering the current and in
the \(x_j\) with hard pion momenta \(p_j\).  Near \(q^2=0\), the current
insertion has the pion pole fixed by
\[
  \bra{\vac}A_a^\mu(0)\ket{\pi^b(q)}
  =
  \ii f_\pi q^\mu\delta^{ab}.
\]
Thus its singular part has the form
\begin{equation}
  \widetilde G_a^\mu(q;\{p_j,a_j\})
  =
  \frac{-\ii f_\pi q^\mu}{q^2-\ii0}\,
  \widetilde G_{n+1}^{\,a,a_1,\ldots,a_n}(q,p_1,\ldots,p_n)
  +\widetilde R_a^\mu(q;\{p_j,a_j\}),
  \label{eq:soft-pion-current-pole-decomposition}
\end{equation}
where \(\widetilde R_a^\mu\) is regular at the current-channel pion pole in the
soft region under consideration.  Let \(\mathcal L_{\{p_j\}}\) denote the hard
external-pion LSZ operation
\[
  \mathcal L_{\{p_j\}}F
  =
  \left[
  \prod_{j=1}^n
  Z_\pi^{-1/2}\,\ii p_j^2
  \right]F
  \quad\hbox{followed by the massless on-shell boundary value }p_j^2=0,
\]
with the same residue convention as in Chapter~\ref{ch:lsz-reduction}.  Applying
\(\mathcal L_{\{p_j\}}\) to the Fourier transform of
\eqref{eq:soft-pion-current-ward-identity} and using
\eqref{eq:soft-pion-current-pole-decomposition} gives
\begin{equation}
  f_\pi\,
  \mathcal M_{n+1}^{a,a_1,\ldots,a_n}(q,p_1,\ldots,p_n)
  =
  \mathcal C_a(q;\{p_j,a_j\})
  +O(q),
  \label{eq:soft-pion-pre-adler-identity}
\end{equation}
where
\[
  \mathcal C_a(q;\{p_j,a_j\})
  =
  \mathcal L_{\{p_j\}}
  \sum_{j=1}^n
  \widetilde G_n^{a_1,\ldots,\delta_A^a a_j,\ldots,a_n}
  (p_1,\ldots,p_j+q,\ldots,p_n)
\]
is the sum of the LSZ-reduced contact terms; the superscript means that the
\(j\)-th pion operator has been replaced by its renormalized axial variation.
This formula is the soft-pion theorem in its current-algebra form: the leading
soft pion is determined by the action of the broken charge on the hard external
states.

\begin{theorem}[Adler zero for chiral pion amplitudes]
\label{thm:adler-zero-chiral-pions}
Assume the exact chiral limit, the symmetry-breaking pattern
\[
  SU(N_f)_L\times SU(N_f)_R\longrightarrow SU(N_f)_V,
\]
the axial-current normalization
\(\bra{\vac}A_a^\mu(0)\ket{\pi^b(p)}=\ii f_\pi p^\mu\delta^{ab}\), and the
existence of the pion scattering boundary values obtained by LSZ from
time-ordered Green functions.  Take the soft momentum \(q\to0\) at fixed
nonexceptional hard momenta, with the hard-leg soft singularities treated by
the LSZ-reduced Ward identity above and no additional threshold or collinear
singular surface crossed.  Then every connected purely pionic amplitude obeys
\begin{equation}
  \mathcal M_{n+1}^{a,a_1,\ldots,a_n}(q,p_1,\ldots,p_n)
  =
  O(q).
  \label{eq:adler-zero-general}
\end{equation}
The vanishing of the \(O(q^0)\) term is the Adler zero.
\end{theorem}

\begin{proof}
The chiral target is the symmetric space
\((SU(N_f)_L\times SU(N_f)_R)/SU(N_f)_V\).  In local pion coordinates normalized
by the axial current, a broken axial transformation has the expansion
\begin{equation}
  \delta_A^a\pi^b
  =
  f_\pi\delta^{ab}
  +O(\pi^2/f_\pi),
  \label{eq:axial-pion-transformation-no-linear-term}
\end{equation}
with no term linear in the pion field.  For \(N_f=2\) this is visible directly
from \eqref{eq:su2-stereographic-chiral-transformation}: using
\(\pi^a=F_{\rm st}\xi^a\) and \(F_{\rm st}=2f_\pi\), the axial variation is
\(\delta_A^a\pi^b=f_\pi\delta^{ab}+O(\pi^2/f_\pi)\).  The absence of a linear
term is the representation-theoretic statement that
\([\mathfrak m,\mathfrak m]\subset\mathfrak h\): a broken generator translates
the origin of the coset and then rotates higher-order tangent data by the
unbroken group, but it does not act as a one-pion-to-one-pion matrix at
nonzero momentum.

Now evaluate the contact term
\(\mathcal C_a(0;\{p_j,a_j\})\) in
\eqref{eq:soft-pion-pre-adler-identity}.  The constant part
\(f_\pi\delta^{aa_j}\) in \eqref{eq:axial-pion-transformation-no-linear-term}
replaces the \(j\)-th interpolating field by the identity operator.  After the
hard LSZ operation it has no one-pion pole in \(p_j^2\), hence gives zero.
The remaining terms in \(\delta_A^a\pi^{a_j}\) contain at least two pion
fields at the same spacetime point.  They are renormalized as part of the same
symmetry-preserving Ward identity, so their projection onto a single hard-pion
pole is fixed to vanish; otherwise the renormalized axial transformation would
have a linear pion term, contrary to
\eqref{eq:axial-pion-transformation-no-linear-term}.  Equivalently, the broken
charge has no one-pion-to-one-pion matrix element; its possible vacuum term is
supported only at zero total momentum and cannot replace a hard external pion.
Therefore the LSZ-reduced contact term has no \(O(q^0)\) contribution:
\[
  \mathcal C_a(0;\{p_j,a_j\})=0 .
\]
Equation \eqref{eq:soft-pion-pre-adler-identity} then gives
\eqref{eq:adler-zero-general}.
\end{proof}

Figure~\ref{fig:soft-pion-ward-adler-zero} records the Ward-identity mechanism
behind the Adler zero: the axial-current insertion is reduced by LSZ, the hard
contact term has no \(O(q^0)\) one-pion contribution, and the soft amplitude
vanishes at leading order.

\begin{figure}[H]
  \centering
  \resizebox{0.93\textwidth}{!}{%
  \begin{tikzpicture}[>=Stealth,x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[draw,rounded corners=2pt,minimum width=2.6cm,minimum height=1.2cm,
      align=center,fill=blue!4] (corr) at (0,0)
      {hard pion\\correlator};
    \draw[decorate,decoration={snake,amplitude=0.7mm,segment length=3.2mm},
      line width=0.8pt,-{Latex[length=2mm]},red!70!black]
      (-3.1,0) -- (corr.west)
      node[midway,above,red!70!black] {\(A_a^\mu(q)\)};
    \draw[thick,-{Latex[length=2mm]}] (corr.east) -- (3.1,0.85)
      node[right] {\(\pi^{a_1}(p_1)\)};
    \draw[thick,-{Latex[length=2mm]}] (corr.east) -- (3.1,0.25)
      node[right] {\(\cdots\)};
    \draw[thick,-{Latex[length=2mm]}] (corr.east) -- (3.1,-0.55)
      node[right] {\(\pi^{a_n}(p_n)\)};
    \node[align=center] at (-2.05,-1.25)
      {current pole\\[-0.2em]
      \(A_a^\mu\sim f_\pi q^\mu\pi^a/q^2\)};
    \node[align=center] at (1.1,-1.32)
      {contact terms\\[-0.2em]
      \(\delta_A^a\pi^b=f_\pi\delta^{ab}+O(\pi^2)\)};
    \draw[-{Latex[length=2mm]},thick] (-0.95,-0.95) -- (0.15,-0.95);
    \node[align=center] at (5.7,0.05)
      {after hard LSZ\\[-0.2em]
      no one-pion-to-one-pion contact\\[-0.2em]
      \(\mathcal M_{n+1}(q,\{p\})=O(q)\)};
    \draw[-{Latex[length=2mm]},thick] (3.35,0.05) -- (4.7,0.05);
  \end{tikzpicture}}
  \caption{The soft-pion theorem follows by isolating the pion pole in the
  axial-current Ward identity.  For purely pionic hard states in the chiral
  coset, the contact terms have no one-pion-to-one-pion component, so the
  leading soft term vanishes.}
  \label{fig:soft-pion-ward-adler-zero}
\end{figure}

For external states that are not all pions, the same identity should be read
before setting \(\mathcal C_a(0)=0\).  A soft pion measures the action of the
broken axial charge on the hard external sector.  If the hard sector contains
degenerate states connected by \(Q_A^a\), the right side of
\eqref{eq:soft-pion-pre-adler-identity} need not vanish.  The Adler zero is the
specialization to the pion \(S\)-matrix in the chiral coset.

For four pions, the theorem says that the analytic on-shell amplitude has a
zero when one external pion momentum is taken soft.  In the chiral limit this
is the statement that the leading invariant amplitude below satisfies
\[
  A(s,t,u)=\frac{s}{f_\pi^2}+O(E^4/f_\pi^4),
  \qquad
  A\longrightarrow0
  \quad\hbox{as}\quad s,t,u\longrightarrow0
\]
in the single-soft analytic continuation.  With a small common pion mass, the
leading chiral formula becomes
\[
  A(s,t,u)=\frac{s-m_\pi^2}{f_\pi^2}+O(p^4),
\]
so the zero is displaced to the corresponding Adler point in the analytically
continued amplitude.  The massive Adler point is not a physical threshold
configuration; it is a Ward-identity zero of the analytic amplitude.

\section{Low-Energy Pion Scattering}

The two-derivative chiral Lagrangian determines the leading low-energy
\(S\)-matrix of the Goldstone bosons, because every interaction is derivative
and therefore ordered by powers of external momentum.  For the \(N_f=2\)
theory it is useful
to write \(SU(2)\simeq S^3\) in stereographic coordinates
\(\vec\xi=(\xi_1,\xi_2,\xi_3)\):
\begin{equation}
  U(\xi)
  =
  \frac{1-\vec\xi^{\,2}+2\ii\,\vec\xi\cdot\vec\sigma}
       {1+\vec\xi^{\,2}} .
  \label{eq:su2-stereographic-pion-field}
\end{equation}
Let \(F_{\rm st}\) denote the normalization used in these stereographic
coordinates:
\begin{equation}
  \mathcal L_2
  =
  -\frac{F_{\rm st}^2}{16}
  \operatorname{Tr}(\partial_\mu U\,\partial^\mu U^\dagger)
  =
  -\frac{F_{\rm st}^2}{2}
  \frac{\partial_\mu\vec\xi\cdot\partial^\mu\vec\xi}
       {(1+\vec\xi^{\,2})^2}.
  \label{eq:su2-stereographic-lagrangian}
\end{equation}
With canonical fields
\[
  \vec\pi=F_{\rm st}\vec\xi,
\]
the expansion begins as
\begin{equation}
  \mathcal L_2
  =
  -\frac12\partial_\mu\vec\pi\cdot\partial^\mu\vec\pi
  +
  \frac1{F_{\rm st}^2}
  \vec\pi^{\,2}\,
  \partial_\mu\vec\pi\cdot\partial^\mu\vec\pi
  +O(\pi^6/F_{\rm st}^4).
  \label{eq:su2-stereographic-expansion}
\end{equation}
The infinitesimal \(SU(2)_L\times SU(2)_R\) action in this chart is useful for
organizing local terms.  Write
\[
  L=\exp\!\left(\frac{\ii}{2}\vec\theta_L\cdot\vec\sigma\right),
  \qquad
  R=\exp\!\left(\frac{\ii}{2}\vec\theta_R\cdot\vec\sigma\right),
  \qquad
  \vec\theta_V=\frac{\vec\theta_L+\vec\theta_R}{2},
  \qquad
  \vec\theta_A=\frac{\vec\theta_L-\vec\theta_R}{2}.
\]
For \(U\mapsto LUR^{-1}\), a direct computation from
\eqref{eq:su2-stereographic-pion-field} gives
\begin{equation}
  \delta\vec\xi
  =
  -\vec\theta_V\times\vec\xi
  +\frac12(1-\vec\xi^{\,2})\vec\theta_A
  +\vec\xi(\vec\xi\cdot\vec\theta_A).
  \label{eq:su2-stereographic-chiral-transformation}
\end{equation}
Thus the diagonal subgroup acts linearly on \(\vec\xi\), while the axial
directions act by a nonlinear shift plus higher powers of \(\vec\xi\).
Define
\begin{equation}
  \vec D_\mu
  =
  \frac{\partial_\mu\vec\xi}{1+\vec\xi^{\,2}} .
  \label{eq:su2-stereographic-d-field}
\end{equation}
For constant transformation parameters,
\begin{equation}
  \delta\vec D_\mu
  =
  -\vec\theta_V\times\vec D_\mu
  -(\vec\xi\times\vec\theta_A)\times\vec D_\mu .
  \label{eq:su2-stereographic-d-transformation}
\end{equation}
The two independent four-derivative invariants relevant to four-pion
scattering can be written as
\[
  (\vec D_\mu\cdot\vec D^\mu)^2,
  \qquad
  (\vec D_\mu\cdot\vec D_\nu)(\vec D^\mu\cdot\vec D^\nu),
\]
up to integrations by parts and terms that first affect higher-point
amplitudes.
This coordinate choice is only a local chart on \(S^3\); the scattering
amplitude obtained from it is coordinate independent.  In the axial-current
normalization above, \(T^a=\sigma^a/2\) and therefore
\[
  \exp\left(\frac{2\ii}{f_\pi}\pi_{\rm ax}^aT^a\right)
  =
  1+\frac{\ii}{f_\pi}\pi_{\rm ax}^a\sigma^a+O(\pi_{\rm ax}^2).
\]
The stereographic chart has
\[
  \frac{1-\vec\xi^{\,2}+2\ii\,\vec\xi\cdot\vec\sigma}
       {1+\vec\xi^{\,2}}
  =
  1+2\ii\,\xi^a\sigma^a+O(\xi^2),
  \qquad
  \pi_{\rm st}^a=F_{\rm st}\xi^a .
\]
Identifying the one-pion field defined by the axial current with the
canonical stereographic field,
\(\pi_{\rm ax}^a=\pi_{\rm st}^a\), gives the conversion
\begin{equation}
  F_{\rm st}=2f_\pi .
  \label{eq:fst-fpi-conversion}
\end{equation}
A convention that rescales the stereographic coordinate rescales
\(F_{\rm st}\) and the coordinate field together; in the fixed chart
\eqref{eq:su2-stereographic-pion-field} the numerical factor in
\eqref{eq:fst-fpi-conversion} is part of the convention.

For
\[
  \pi^a(k_1)+\pi^b(k_2)\to \pi^c(k_3)+\pi^d(k_4),
\]
use the Mandelstam variables
\[
  s=-(k_1+k_2)^2,\qquad
  t=-(k_1-k_3)^2,\qquad
  u=-(k_1-k_4)^2 .
\]
The metric is mostly plus, \(k_1,k_2\) are incoming, and \(k_3,k_4\) are
outgoing.  In the chiral limit used for the two-derivative Goldstone
calculation, \(k_i^2=0\), so
\[
  s+t+u=0 .
\]
For equal nonzero pion masses the same convention gives
\(s+t+u=4m_\pi^2\).
The \(SU(2)_V\)-invariant amplitude has the form
\begin{equation}
  \mathcal A_{ab\to cd}
  =
  \delta_{ab}\delta_{cd}A(s,t,u)
  +\delta_{ac}\delta_{bd}A(u,t,s)
  +\delta_{ad}\delta_{bc}A(t,s,u).
  \label{eq:pion-scattering-isospin-decomposition}
\end{equation}
At order \(E^2/F_{\rm st}^2\), the four-pion vertex from
\eqref{eq:su2-stereographic-expansion} gives
\begin{equation}
  A(s,t,u)
  =
  \frac{4s}{F_{\rm st}^2}
  +O(E^4/F_{\rm st}^4).
  \label{eq:pion-leading-scattering-amplitude}
\end{equation}
Equivalently, after converting to the convention in which the axial current
matrix element defines \(f_\pi\), this is the standard leading chiral theorem
\[
  A(s,t,u)=\frac{s}{f_\pi^2}+O(E^4/f_\pi^4).
\]

Figure~\ref{fig:pion-scattering-leading-next} displays the two levels of the
chiral expansion used here: the leading two-derivative four-pion vertex and the
next-order loop/local-counterterm structure.

\begin{figure}[H]
  \centering
  \resizebox{0.9\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small] at (0,2.0) {leading two-derivative vertex};
      \draw[thick] (-1.7,1.0)--(0,0)--(1.7,1.0);
      \draw[thick] (-1.7,-1.0)--(0,0)--(1.7,-1.0);
      \fill (0,0) circle (2.2pt);
      \node[left] at (-1.7,1.0) {\(a,k_1\)};
      \node[left] at (-1.7,-1.0) {\(b,k_2\)};
      \node[right] at (1.7,1.0) {\(c,k_3\)};
      \node[right] at (1.7,-1.0) {\(d,k_4\)};
      \node[below] at (0,-1.42) {\(A(s,t,u)=4s/F_{\rm st}^2\)};
    \end{scope}

    \begin{scope}[shift={(6.2,0)}]
      \node[font=\small] at (0,2.0) {next order in the chiral expansion};
      \draw[thick] (-1.95,0.65)--(-0.75,0.24);
      \draw[thick] (-1.95,-0.65)--(-0.75,-0.24);
      \draw[thick] (0.75,0.24)--(1.95,0.65);
      \draw[thick] (0.75,-0.24)--(1.95,-0.65);
      \draw[thick] (-0.75,0.24) arc (160:-160:0.80);
      \draw[thick] (0.75,-0.24) arc (-20:200:0.80);
      \node at (2.48,0) {\(+\)};
      \begin{scope}[shift={(4.45,0)}]
        \draw[thick] (-1.25,0.75)--(0,0)--(1.25,0.75);
        \draw[thick] (-1.25,-0.75)--(0,0)--(1.25,-0.75);
        \fill (0,0) circle (2.2pt);
        \node[below] at (0,-0.35) {\(C_4,C'_4\)};
      \end{scope}
      \node[align=center] at (1.9,-1.42)
        {one loop from \(\mathcal L_2\)\\and local four-derivative terms};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The leading pion scattering amplitude is fixed by the
  two-derivative nonlinear sigma model.  At the next order, one-loop diagrams
  from \(\mathcal L_2\) are combined with local four-derivative counterterms.}
  \label{fig:pion-scattering-leading-next}
\end{figure}

At order \(E^4/F_{\rm st}^4\), the effective Lagrangian contains local
four-derivative terms.  In the stereographic \(SU(2)\) chart these may be
written, to the order needed for four-pion scattering, as
\[
  -\frac{C_4}{4F_{\rm st}^4}
  (\partial_\mu\vec\pi\cdot\partial^\mu\vec\pi)^2
  -
  \frac{C'_4}{4F_{\rm st}^4}
  (\partial_\mu\vec\pi\cdot\partial_\nu\vec\pi)
  (\partial^\mu\vec\pi\cdot\partial^\nu\vec\pi).
\]
The same order also receives one-loop contributions from two insertions of
the two-derivative interaction.  A regulator that does not preserve the
nonlinearly realized symmetry may generate power terms that look like a
potential on \(S^3\); such terms must be absent in a symmetry-preserving
scheme or canceled by symmetry-restoring local counterterms.  After this
local renormalization, the logarithmic part has the form
\begin{align}
  A(s,t,u)
  &=
  \frac{4s}{F_{\rm st}^2}
  +
  \frac1{F_{\rm st}^4}
  \biggl\{
  \frac1{12\pi^2}
    \biggl[
      3s^2\log\frac{\mu^2}{-s}
      \notag\\
  &\hspace{2.4cm}
      +(u^2-s^2+3t^2)\log\frac{\mu^2}{-t}
      \notag\\
  &\hspace{2.4cm}
      +(t^2-s^2+3u^2)\log\frac{\mu^2}{-u}
    \biggr]
    \notag\\
  &\hspace{1.6cm}
    -\frac12 C_4(\mu)s^2
    -\frac12 C'_4(\mu)(t^2+u^2)
  \biggr\}
  +O(E^6).
  \label{eq:pion-one-loop-chiral-log-amplitude}
\end{align}
Equivalently, in a Wilsonian cutoff notation in which the logarithms are
written as \(\log(\Lambda^2/-s)\), \(\log(\Lambda^2/-t)\), and
\(\log(\Lambda^2/-u)\), the remaining cutoff dependence is absorbed by
\[
  C_4(\Lambda)
  =
  C_4(\mu)+\frac{2}{3\pi^2}\log\frac{\Lambda^2}{\mu^2},
  \qquad
  C'_4(\Lambda)
  =
  C'_4(\mu)+\frac{4}{3\pi^2}\log\frac{\Lambda^2}{\mu^2}.
\]
The finite constants \(C_4(\mu)\) and \(C'_4(\mu)\) are low-energy constants.
Their values depend on the microscopic dynamics and on the subtraction
convention, but the logarithmic dependence and the required \(\mu\)-dependence
are fixed by the low-energy effective theory.

\section{Mass Spurions and Pseudo-Goldstone Bosons}

Quark masses explicitly break the chiral symmetry.  Introduce a complex mass
matrix \(M\) through
\[
  \mathcal L_{\mathrm{mass}}^{\mathrm{QCD}}
  =
  -\bar q_L M q_R-\bar q_RM^\dagger q_L .
\]
Treating \(M\) as a background spurion with
\[
  M\longmapsto LMR^{-1}
\]
makes the microscopic mass term formally invariant under
\((L,R)\).  The leading mass-dependent invariant in the pion Lagrangian is
\begin{equation}
  \mathcal L_{\mathrm{mass}}
  =
  \frac{f_\pi^2}{4}
  \operatorname{Tr}
  \bigl(\chi U^\dagger+U\chi^\dagger\bigr),
  \qquad
  \chi=2B_0M .
  \label{eq:chiral-mass-spurion}
\end{equation}
The low-energy constant \(B_0\) is fixed by the chiral-limit condensate
convention
\[
  \langle\bar q q\rangle_0
  =
  -N_f f_\pi^2B_0
\]
at leading order.  For two light flavors,
\[
  m_{\pi^\pm}^2
  =
  B_0(m_u+m_d)
\]
up to higher-order chiral and electromagnetic corrections.  For degenerate
light quarks \(m_u=m_d=m_q\), this becomes \(m_\pi^2=2B_0m_q\).
The condensate notation in this chapter is flavor-summed:
\[
  \Sigma_{\mathrm{sum}}
  :=
  \langle\bar q q\rangle_0
  =
  \sum_{i=1}^{N_f}\langle\bar q_iq_i\rangle_0,
  \qquad
  \Sigma_{\mathrm{fl}}
  :=
  \frac{1}{N_f}\Sigma_{\mathrm{sum}} .
\]
The leading-order identity following from the two preceding equations is the
flavor-summed form
\begin{equation}
  m_{\pi^\pm}^2f_\pi^2
  =
  -(m_u+m_d)\,
  \frac{\langle\bar q q\rangle_0}{N_f}.
  \label{eq:gmor-flavor-summed}
\end{equation}
The conversion to the per-flavor condensate is
\begin{equation}
  \Sigma_{\mathrm{fl}}
  =
  \frac{\langle\bar q q\rangle_0}{N_f}
  =
  \langle\bar u u\rangle_0
  =
  \langle\bar d d\rangle_0
  \qquad
  (N_f=2\hbox{ isospin-symmetric vacuum}).
  \label{eq:gmor-condensate-conversion}
\end{equation}
Using \eqref{eq:gmor-condensate-conversion}, the same leading-order statement is
the per-flavor Gell-Mann--Oakes--Renner relation
\begin{equation}
  m_{\pi}^2f_\pi^2
  =
  -(m_u+m_d)\langle\bar u u\rangle_0 .
  \label{eq:gmor-per-flavor}
\end{equation}
The factor of \(2\) is accounted for by the conversion between the
flavor-summed condensate and the per-flavor condensate.

The same normalization can be seen directly in the \(SU(2)\) stereographic
chart.  Let
\[
  M=\begin{pmatrix}m_u&0\\0&m_d\end{pmatrix}.
\]
In terms of the quark bilinear \(\Phi^I{}_{J}=\bar q_L^Iq_{R,J}\), the
microscopic deformation has the form
\[
  \Delta\mathcal L
  =
  -\frac{m_u+m_d}{2}
  \operatorname{Tr}(\Phi+\Phi^\dagger)
  -\frac{m_u-m_d}{2}
  \operatorname{Tr}\bigl(\sigma_3(\Phi+\Phi^\dagger)\bigr).
\]
Restricting the pion field to the \(U(1)\subset SU(2)\) generated by
\(\sigma_3\),
\[
  U(\xi_3)
  =
  1+2\ii\xi_3\sigma_3+\cdots
  =
  \begin{pmatrix}\ee^{\ii\theta}&0\\0&\ee^{-\ii\theta}\end{pmatrix}
  +\cdots ,
  \qquad
  \theta=2\xi_3+\cdots .
\]
Since the kinetic term in \eqref{eq:su2-stereographic-lagrangian} gives
\[
  \mathcal L_2=-\frac{F_{\rm st}^2}{2}\partial_\mu\vec\xi\cdot
  \partial^\mu\vec\xi+\cdots,
\]
the \(U(1)\) comparison has radius \(R=F_{\rm st}/2\).  Expanding the mass
deformation therefore gives
\[
  -\frac12 m_{ab}^2\pi_a\pi_b,
  \qquad
  m_{ab}^2
  =
  \delta_{ab}\frac{4}{F_{\rm st}^2}(m_u+m_d)
  \left\langle-\frac{\bar q q}{2}\right\rangle_0 ,
\]
which is the same leading relation as the \(B_0\)-notation after converting
normalizations.  Empirically,
\[
  m_{\pi^0}\simeq135\,\mathrm{MeV},
  \qquad
  m_{\pi^\pm}\simeq140\,\mathrm{MeV},
\]
with the splitting produced by isospin breaking and electromagnetism.

The chiral expansion counts \(D_\mu U=O(p)\), external vector and axial
backgrounds as \(O(p)\), and \(M\) as \(O(p^2)\).  The functional obtained from
\(\mathcal L_2+\mathcal L_{\mathrm{mass}}\) is therefore the leading
generating functional for current Green functions at momenta small compared
with the chiral-symmetry-breaking scale.

\section{Chiral Perturbation Theory at NLO}
\label{sec:chiral-perturbation-theory-nlo}

Chiral perturbation theory is the effective expansion of the QCD current
generating functional in powers of derivatives, external flavor fields, and
quark masses around the chiral limit.  Pion fields enter as local coordinates
on the Goldstone manifold; the invariant expansion is organized by the orders
of the functional's derivatives, sources, and spurion insertions.  The
expansion variables are assigned the orders
\[
  D_\mu U=O(p),\qquad
  \ell_\mu,r_\mu=O(p),\qquad
  F^L_{\mu\nu},F^R_{\mu\nu}=O(p^2),\qquad
  \chi=2B_0M=O(p^2),
\]
where the left and right field strengths in the conventions of
\eqref{eq:pion-external-covariant-derivative} are
\[
  F^L_{\mu\nu}
  =
  \partial_\mu\ell_\nu-\partial_\nu\ell_\mu
  -\ii[\ell_\mu,\ell_\nu],
  \qquad
  F^R_{\mu\nu}
  =
  \partial_\mu r_\nu-\partial_\nu r_\mu
  -\ii[r_\mu,r_\nu].
\]
A connected graph built from vertices of chiral order \(d_i\) and with \(L\)
loops has total chiral order
\begin{equation}
  \nu
  =
  2+2L+\sum_i V_i(d_i-2).
  \label{eq:weinberg-chiral-power-counting}
\end{equation}
Indeed, each loop integration contributes \(p^4\), each internal pion
propagator contributes \(p^{-2}\), and each vertex contributes \(p^{d_i}\).
For a connected graph, \(L=I-\sum_iV_i+1\), hence
\[
  4L-2I+\sum_i V_id_i
  =
  2+2L+\sum_iV_i(d_i-2).
\]
Thus one-loop graphs made only from \(\mathcal L_2\) and tree graphs from the
order-\(p^4\) Lagrangian both contribute at next-to-leading order.

\begin{controlledapproximation}[Chiral EFT truncation as a seminorm statement]
\label{ca:chiral-eft-truncation-seminorm}
The power-counting identity \eqref{eq:weinberg-chiral-power-counting} is only
the graph algebra.  To use it as a QCD statement one must also specify the
observable topology in which the low-energy remainder is controlled.  Fix a
compact family \(K\) of external momenta and smooth flavor sources satisfying
\[
  |p_a|\le \epsilon\Lambda_\chi,\qquad
  \|\ell\|_K+\|r\|_K\le \epsilon\Lambda_\chi,\qquad
  \|\chi\|_K^{1/2}\le \epsilon\Lambda_\chi,
  \qquad 0<\epsilon<1,
\]
together with a renormalization scheme for the local chiral coordinates.  For
a connected renormalized graph \(\Gamma\) let
\[
  \nu(\Gamma)=2+2L(\Gamma)+\sum_d V_d(\Gamma)(d-2).
\]
The useful truncation assertion is a family of seminorm estimates
\[
  p_K\!\left(\mathcal A_\Gamma\right)
  \le C_{\Gamma,K}\epsilon^{\nu(\Gamma)}
\]
for the source derivatives or \(S\)-matrix coordinates under consideration,
and, after retaining all graphs with \(\nu\le N\),
\begin{equation}
  p_K\!\left(
    \mathcal A-\sum_{\nu(\Gamma)\le N}\mathcal A_\Gamma
  \right)
  \le C_{K,N}\epsilon^{\nu_+(N)},
  \qquad
  \nu_+(N):=\min\{\nu(\Gamma):\nu(\Gamma)>N\}.
  \label{eq:chiral-eft-seminorm-remainder}
\end{equation}
Equation \eqref{eq:chiral-eft-seminorm-remainder} is not a convergence theorem
for the full chiral series.  It is the controlled EFT input that turns the
formal graph dimension into a quantitative low-energy approximation on the
declared compact source window.  At \(N=4\) the retained data are precisely:
tree graphs from \(\mathcal L_2\), one-loop graphs from \(\mathcal L_2\), and
tree graphs from \(\mathcal L_4\), including the source-contact terms.  The
first omitted families have \(\nu=6\): two-loop graphs from \(\mathcal L_2\),
one-loop graphs with one \(\mathcal L_4\) vertex, and tree graphs from
\(\mathcal L_6\).  Thus the \(O(p^4)\) calculation below is meaningful only
inside this remainder ledger; outside such a low-energy window the same local
formulae are just coordinates, not controlled predictions.
\end{controlledapproximation}

For \(N_f=3\), the even-intrinsic-parity order-\(p^4\) local functional is the
Gasser--Leutwyler basis.\footnote{The notation \(L_1,\ldots,L_{10}\) is the
three-flavor basis introduced by Gasser and Leutwyler in the original chiral
perturbation theory papers, translated here to the left/right convention
\(U\mapsto LUR^{-1}\).}  Write \(\langle X\rangle_{\rm f}\) for the flavor
trace \(\operatorname{Tr}X\), and define
\[
  X_+ := \chi U^\dagger+U\chi^\dagger,
  \qquad
  X_- := \chi U^\dagger-U\chi^\dagger .
\]
With the left/right convention \(U\mapsto LUR^{-1}\), the local NLO
Lagrangian is
\begin{align}
  \mathcal L_4^{\rm GL}
  &=
  L_1\langle D_\mu U^\dagger D^\mu U\rangle_{\rm f}^2
  +
  L_2\langle D_\mu U^\dagger D_\nu U\rangle_{\rm f}
      \langle D^\mu U^\dagger D^\nu U\rangle_{\rm f}
  \notag\\
  &\quad
  +
  L_3\langle
    D_\mu U^\dagger D^\mu U
    D_\nu U^\dagger D^\nu U
  \rangle_{\rm f}
  +
  L_4\langle D_\mu U^\dagger D^\mu U\rangle_{\rm f}
      \langle X_+\rangle_{\rm f}
  \notag\\
  &\quad
  +
  L_5\langle D_\mu U D^\mu U^\dagger X_+\rangle_{\rm f}
  +
  L_6\langle X_+\rangle_{\rm f}^2
  +
  L_7\langle X_-\rangle_{\rm f}^2
  \notag\\
  &\quad
  +
  L_8\langle
    \chi U^\dagger\chi U^\dagger
    +U\chi^\dagger U\chi^\dagger
  \rangle_{\rm f}
  \notag\\
  &\quad
  -\ii L_9
  \langle
    F_L^{\mu\nu}D_\mu U D_\nu U^\dagger
    +F_R^{\mu\nu}D_\mu U^\dagger D_\nu U
  \rangle_{\rm f}
  \notag\\
  &\quad
  +
  L_{10}
  \langle U^\dagger F_L^{\mu\nu}U F^R_{\mu\nu}\rangle_{\rm f}
  \notag\\
  &\quad
  +
  H_1\langle F_L^{\mu\nu}F^L_{\mu\nu}
      +F_R^{\mu\nu}F^R_{\mu\nu}\rangle_{\rm f}
  +
  H_2\langle \chi^\dagger\chi\rangle_{\rm f}.
  \label{eq:gasser-leutwyler-l1-l10-basis}
\end{align}
The constants \(L_1,\ldots,L_{10}\) multiply operators containing the pion
field or its covariant derivatives.  The contact constants \(H_1,H_2\) affect
only correlators of external sources and are required for renormalizing the
source generating functional.  The odd-intrinsic-parity anomaly sector is not
part of \(\mathcal L_4^{\rm GL}\); it is the Wess--Zumino--Witten functional
constructed in the next section.  For \(N_f=2\), Cayley--Hamilton identities
reduce the number of independent even-parity \(p^4\) invariants.  The
stereographic constants \(C_4,C'_4\) above are two-coordinate representatives
of that reduced two-flavor sector; the \(L_i\) above are the standard
three-flavor coordinates.

The \(L_i\) are renormalized coordinates on the space of local order-\(p^4\)
counterterms.  In dimensional regularization one writes
\[
  L_i=L_i^r(\mu)+\Gamma_i\Lambda_{\rm div},
  \qquad
  \mu\frac{\dd L_i^r}{\dd\mu}
  =
  -\frac{\Gamma_i}{16\pi^2},
\]
where the three-flavor coefficients are
\[
\begin{array}{c|cccccccccc}
i&1&2&3&4&5&6&7&8&9&10\\ \hline
\Gamma_i&
\frac{3}{32}&\frac{3}{16}&0&\frac18&\frac38&
\frac{11}{144}&0&\frac{5}{48}&\frac14&-\frac14 .
\end{array}
\]
The \(\mu\)-dependence of \(L_i^r(\mu)\) cancels the logarithmic dependence
from one-loop graphs built from \(\mathcal L_2\).  This cancellation is a
low-energy theorem; the finite values of \(L_i^r(\mu)\) encode microscopic
QCD dynamics.

The simplest chiral logarithm appears in the two-flavor pion mass.  Set
\(m_u=m_d=m_q\) and
\[
  M^2:=2B_0m_q .
\]
The one-loop tadpole integral generated by the \(\mathcal L_2\) four-pion
vertices has renormalized logarithmic part
\[
  A^r(M^2)
  =
  \frac{M^2}{16\pi^2}\log\frac{M^2}{\mu^2}.
\]
Expanding \(\mathcal L_2\) in normal pion coordinates gives the logarithmic
on-shell self-energy contribution
\[
  \Sigma_{\log}(M^2)
  =
  \frac{M^2}{2f^2}A^r(M^2)
  =
  \frac{M^4}{32\pi^2f^2}
  \log\frac{M^2}{\mu^2},
\]
where \(f\) is the chiral-limit value of \(f_\pi\).  The local two-flavor NLO
constant usually denoted \(l_3^r(\mu)\) contributes \(2l_3^r(\mu)M^4/f^2\) to
the mass.  Hence
\begin{equation}
  M_{\pi,\rm phys}^2
  =
  M^2
  \left[
    1
    +
    \frac{M^2}{32\pi^2f^2}\log\frac{M^2}{\mu^2}
    +
    \frac{2l_3^r(\mu)M^2}{f^2}
  \right]
  +O(M^6).
  \label{eq:su2-pion-mass-nlo-chiral-log}
\end{equation}
This is nonanalytic in the quark mass because \(M^2\propto m_q\).  The
renormalization-group coefficient of the two-flavor constant is
\[
  \mu\frac{\dd l_3^r}{\dd\mu}
  =
  \frac{1}{32\pi^2},
\]
so differentiating the bracket in
\eqref{eq:su2-pion-mass-nlo-chiral-log} with respect to \(\log\mu\) gives
\[
  -\frac{M^2}{16\pi^2f^2}
  +
  \frac{2M^2}{f^2}\frac{1}{32\pi^2}
  =
  0.
\]
Thus the scale dependence of the logarithm is exactly canceled by the running
of the local NLO coordinate.  The logarithm itself cannot be removed by a
local counterterm because it is the long-distance contribution of pion
propagation.

\section{The Wess--Zumino--Witten Functional}

The pion effective action must reproduce the global anomalies of the
microscopic quarks.  For \(N_f\ge3\), let \(M_4\) be spacetime and choose a
five-manifold \(B_5\) with \(\partial B_5=M_4\).  Extend
\(U:M_4\to SU(N_f)\) to \(\widetilde U:B_5\to SU(N_f)\).
Figure~\ref{fig:wzw-five-dimensional-filling} fixes this filling convention for
the Wess--Zumino--Witten functional: \(M_4\) is the boundary and
\(\widetilde U\) is an extension over \(B_5\).
\begin{figure}[H]
  \centering
  \begin{tikzpicture}[>=Stealth,x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \draw[thick] (0,0) ellipse (2.0 and 1.25);
    \draw[thick] (-2,0) arc (180:360:2.0 and 0.42);
    \draw[dashed] (2,0) arc (0:180:2.0 and 0.42);
    \fill[blue!12] (-1.58,0.08)
      .. controls (-0.9,0.42) and (0.25,0.35) .. (1.18,0.12)
      .. controls (0.8,-0.12) and (-0.5,-0.18) .. (-1.58,0.08);
    \draw[blue!70,thick] (-1.58,0.08)
      .. controls (-0.9,0.42) and (0.25,0.35) .. (1.18,0.12)
      .. controls (0.8,-0.12) and (-0.5,-0.18) .. (-1.58,0.08);
    \node[blue!70] at (-0.15,0.15) {\(B_5\)};
    \node at (2.25,0.35) {\(M_4=\partial B_5\)};
    \draw[->] (1.95,0.28) -- (1.35,0.1);
  \end{tikzpicture}
  \caption{The Wess--Zumino functional is written by extending the spacetime
  field \(U:M_4\to SU(N_f)\) to a five-dimensional filling \(B_5\).  Changing
  the filling changes the integral by a quantized period.}
  \label{fig:wzw-five-dimensional-filling}
\end{figure}
\begin{definition}[Ungauged Wess--Zumino functional]
\label{def:wess-zumino-witten-functional}
The ungauged Wess--Zumino functional is
\begin{equation}
  \Gamma_{\mathrm{WZ}}[U]
  =
  -\frac{\ii}{240\pi^2}
  \int_{B_5}
  \operatorname{Tr}
  \bigl(\widetilde U^{-1}\dd \widetilde U\bigr)^5 .
  \label{eq:wess-zumino-functional}
\end{equation}
The five-form is real after the factor \(-\ii\), because
\(\widetilde U^{-1}\dd\widetilde U\) is anti-Hermitian.  If two extensions are
glued along \(M_4\), the difference of \(\Gamma_{\mathrm{WZ}}\) is the
integral over a closed five-manifold.  With the normalization in
\eqref{eq:wess-zumino-functional}, this difference is \(2\pi\) times an
integer for maps representing \(\pi_5(SU(N_f))\simeq\mathbb Z\).  Hence
\[
  \exp\{\ii n\Gamma_{\mathrm{WZ}}[U]\}
\]
is independent of the extension precisely when \(n\in\mathbb Z\).  This
integer is the Wess--Zumino level.
\end{definition}

\begin{proposition}[Finite Wess--Zumino extension ambiguity]
\label{prop:finite-wz-extension-ambiguity}
For \(N_f\ge3\), write
\[
  \omega_5(U)
  =
  -\frac{\ii}{240\pi^2}
  \operatorname{Tr}(U^{-1}\dd U)^5 .
\]
Assume the standard generator normalization in which
\(\int_{S^5}g^*\omega_5=2\pi\) for a generator
\([g]\in\pi_5(SU(N_f))\simeq\mathbb Z\).  If two fillings
\((B_5,\widetilde U)\) and \((B'_5,\widetilde U')\) extend the same boundary
field \(U:M_4\to SU(N_f)\), then
\[
  \Gamma_{\mathrm{WZ}}[B_5,\widetilde U]
  -
  \Gamma_{\mathrm{WZ}}[B'_5,\widetilde U']
  =
  2\pi k,\qquad k\in\mathbb Z .
\]
Consequently the exponentiated pion response
\(\exp\{\ii n\Gamma_{\mathrm{WZ}}[U]\}\) is a well-defined boundary
functional for all allowed fillings exactly when \(n\in\mathbb Z\).
\end{proposition}

\begin{proof}
Glue \(B_5\) to \(B'_5\) with the opposite orientation along their common
boundary \(M_4\).  The result is a closed oriented five-manifold
\[
  X_5=B_5\cup_{M_4}(-B'_5),
\]
and the two extensions glue to a map \(\widehat U:X_5\to SU(N_f)\).
Additivity of the integral under gluing gives
\[
  \Gamma_{\mathrm{WZ}}[B_5,\widetilde U]
  -
  \Gamma_{\mathrm{WZ}}[B'_5,\widetilde U']
  =
  \int_{X_5}\widehat U^*\omega_5 .
\]
The normalized period assumption says that the right hand side lies in
\(2\pi\mathbb Z\); call it \(2\pi k\).  Changing the filling therefore
multiplies the exponentiated level-\(n\) response by
\[
  \exp(2\pi\ii nk).
\]
This factor is \(1\) for every possible extension ambiguity \(k\in\mathbb Z\)
if \(n\) is an integer.  Conversely, the generator ambiguity \(k=1\) would
give a nontrivial phase for any nonintegral level.  Thus finite
extension-independence is the integrality condition \(n\in\mathbb Z\).
\end{proof}

The finite integrality statement is separate from the infinitesimal anomaly
coefficient comparison below.  Local descent determines which multiple of the
generator the pion functional must carry; the proposition says when that
multiple exponentiates to an honest four-dimensional response rather than a
choice of five-dimensional filling.  In QCD the same integer is \(N_c\), the
number of color copies of a fundamental flavor Weyl fermion.

\paragraph{WZW level matching from the flavor descent coefficient.}
\label{par:wzw-anomaly-matching}
The equality between the QCD Wess--Zumino--Witten level and the number of
colors is not an additional imported theorem.  It is the coefficient matching
of two representatives of the same infinitesimal background-anomaly class,
after the flavor trace convention
\(\operatorname{Tr}(T^aT^b)=\frac12\delta^{ab}\) and the unit-level
normalization of \(\Gamma_{\mathrm{WZ}}\) in
\eqref{eq:wess-zumino-functional} have been fixed.  For massless QCD with
\(N_c\) colors and \(N_f\ge3\) light flavors, gauge a left flavor background
\(A=A_\mu\dd x^\mu\), set the right background to zero, and first retain the
part quadratic in \(A\).  Matrix multiplication and wedge product are
implicit in the following formulas.

The word ``quadratic'' here does not mean that the higher-background part of
the anomaly is being discarded.  The previous chapter showed that a local
Bardeen counterterm can change a four-form representative \(I_4^{(1)}\), but
cannot change the completely symmetric degree-three coefficient of the
six-form polynomial \(I_6\).  More explicitly, the Abelianized descent
coordinate \eqref{eq:abelianized-cubic-descent-coordinate} proves that, on
commuting flavor backgrounds, the boundary variation of the Chern--Simons
representative is exactly \(C_{abc}\alpha^aF^bF^c\), and the polarization
identity \eqref{eq:cubic-polarization-identity-for-anomaly-coordinate}
recovers the symmetric cubic coefficient from these commuting tests.  The term
\(\operatorname{Tr}(\alpha\,\dd A\,\dd A)\) is the lowest-background
coordinate of that same descent class in the convention fixed in
\eqref{eq:consistent-descent-anomaly-representative}.  The \(O(A^3)\) terms
are the non-Abelian completion required by Wess--Zumino consistency inside
the same class, up to local counterterm coboundaries.  A single left-handed
quark multiplet gives the consistent variation
\[
  \delta_\alpha\Gamma_{\mathrm{1\,color}}[A]
  =
  \frac1{48\pi^2}
  \int_{M_4}\operatorname{Tr}
  \bigl(\alpha\,\dd A\,\dd A\bigr)
  +O(A^3),
\]
where \(\alpha=\alpha^aT^a\).  QCD has \(N_c\) identical color copies of this
flavor representation, hence
\[
  \delta_\alpha\Gamma_{\mathrm{quark}}[A]
  =
  \frac{N_c}{48\pi^2}
  \int_{M_4}\operatorname{Tr}
  \bigl(\alpha\,\dd A\,\dd A\bigr)
  +O(A^3).
\]
The gauged Wess--Zumino functional at level \(n\) has the corresponding
Goldstone variation
\[
  \delta_\alpha\bigl(n\Gamma_{\mathrm{WZW}}[U;A,0]\bigr)
  =
  \frac{n}{48\pi^2}
  \int_{M_4}\operatorname{Tr}
  \bigl(\alpha\,\dd A\,\dd A\bigr)
  +O(A^3).
\]
Equality of the anomaly phases for arbitrary infinitesimal \(\alpha\) and
background \(A\) gives
\[
  n=N_c .
\]
The comparison of the displayed quadratic term fixes the integer coefficient
of the generator of the descent class.  This is why the level is fixed by
one coefficient rather than by a term-by-term comparison of a chosen
representative.  The remaining background terms are determined by the same
relative BRST class after the counterterm convention has been chosen.  When
the flavor symmetry is coupled to background gauge fields, the Wess--Zumino
functional is completed by local four-dimensional terms involving
\(U,\ell_\mu,r_\mu\).  The completed Wess--Zumino--Witten functional
\(\Gamma_{\mathrm{WZW}}[U;\ell,r]\) satisfies
\[
  \delta_{\alpha_L,\alpha_R}
  \Gamma_{\mathrm{WZW}}[U;\ell,r]
  =
  \mathcal A_{\mathrm{quark}}[\alpha_L,\alpha_R;\ell,r],
\]
where \(\mathcal A_{\mathrm{quark}}\) is the consistent flavor anomaly of the
microscopic quarks.  Thus the pion functional supplies exactly the anomaly
phase required by \eqref{eq:finite-background-anomaly}.  Equivalently, the
gauged Wess--Zumino--Witten functional is a Goldstone-field realization of the
same inflow class: its five-dimensional extension transgresses the microscopic
anomaly polynomial, and its four-dimensional gauge variation is the boundary
term that matches the quark anomaly.

\section{'t Hooft Anomaly Matching as an Infrared Constraint}
\label{sec:t-hooft-anomaly-matching-tool}

Anomaly matching becomes a calculational constraint only after the symmetry,
backgrounds, and counterterm convention have been fixed.  The procedure is:
\begin{enumerate}[label=\roman*.]
\item choose the exact global symmetry that is being tested and the class of
allowed background fields;
\item compute the ultraviolet background-gauge cocycle, or equivalently its
infinitesimal Wess--Zumino representative on the identity component;
\item write the most general proposed infrared response from massless
fermions, Goldstone fields, and possible topological or invertible sectors;
\item compare cocycle classes modulo local counterterms.
\end{enumerate}
The comparison is not a comparison of Lagrangian terms or of a particular
representative of a current.  It is a comparison of the background response
class \([\mathcal A]\) defined in \eqref{eq:background-anomaly-cocycle}.  A
candidate infrared phase with a unique gapped vacuum and no topological
response has the zero class.  Such a phase is ruled out when the ultraviolet
class is nonzero.  A candidate phase with Goldstone fields is allowed only if
its Wess--Zumino--Witten term supplies the missing cocycle.

\begin{remark}[Anomaly matching across spontaneous symmetry breaking]
\label{rem:anomaly-matching-across-ssb}
Suppose an exact global symmetry \(G_{\mathrm f}\) is spontaneously broken to
\(H\), and the infrared fields consist of Goldstone coordinates on \(G/H\)
together with possible additional massless or topological sectors.  Let
\([\mathcal A_{\rm UV}]\) be the ultraviolet background-anomaly class of
Definition~\ref{def:background-anomaly-cocycle}.  The infrared description is
compatible with the ultraviolet theory only if
\[
  [\mathcal A_{\rm UV}]
  =
  [\mathcal A_{\rm Goldstone}]
  +
  [\mathcal A_{\rm extra}]
\]
as background-response classes modulo local counterterms.  Here
\([\mathcal A_{\rm Goldstone}]\) is produced by the gauged WZW functional
whenever such a functional exists, and \([\mathcal A_{\rm extra}]\) is the
class carried by the remaining infrared sector.

Along an RG trajectory with exact \(G_{\mathrm f}\) background fields, the
finite-regulator shell pushforward in
Remark~\ref{rem:anomaly-matching-background-cocycle-class} transports the same
anomaly line.  This is the regulator-level statement; the continuum and
long-distance endpoint comparison requires the compatibility assumptions
stated there, and is not derived from a local background-only ratio of endpoint
partition functions.  In the broken phase, the infrared section is supplied by
the Goldstone Wess--Zumino--Witten functional, the remaining massless or
topological sector, and possible local retained-field and background
counterterms.  Under a background gauge transformation, their nonlocal pieces
contribute the sum of their cocycle representatives, while local terms shift
representatives by coboundaries.  Passing to background-response classes
therefore gives the displayed equality.
\end{remark}

For massless QCD, restrict first to the connected nonabelian chiral symmetry
\[
  SU(N_f)_L\times SU(N_f)_R ,
  \qquad N_f\ge3,
\]
and use the flavor generators \(T^a\) normalized by
\(\operatorname{Tr}(T^aT^b)=\frac12\delta^{ab}\), as in the pion
normalization above.  Couple only a left background \(A\) and set the right
background to zero.  The ultraviolet theory contains \(N_c\) color copies of a
left-handed Weyl fermion in the fundamental representation of
\(SU(N_f)_L\).  The consistent anomaly representative is therefore
\begin{equation}
  \delta_\alpha W_{\rm UV}[A]
  =
  \frac{N_c}{48\pi^2}
  \int_{M_4}
  \operatorname{Tr}\bigl(\alpha\,\dd A\,\dd A\bigr)
  +O(A^3),
  \label{eq:qcd-left-flavor-uv-anomaly-repeat}
\end{equation}
where the \(O(A^3)\) terms are fixed by the same descent class.  A
two-derivative pion sigma model without a Wess--Zumino--Witten term is
invariant under background gauge transformations and would give zero on the
right hand side of \eqref{eq:qcd-left-flavor-uv-anomaly-repeat}; it therefore
cannot be the complete infrared response to the chiral backgrounds.  A
Goldstone theory with WZW level \(n\) gives
\[
  \delta_\alpha\bigl(n\Gamma_{\mathrm{WZW}}[U;A,0]\bigr)
  =
  \frac{n}{48\pi^2}
  \int_{M_4}
  \operatorname{Tr}\bigl(\alpha\,\dd A\,\dd A\bigr)
  +O(A^3).
\]
Matching \eqref{eq:qcd-left-flavor-uv-anomaly-repeat} for arbitrary
\(\alpha\) and \(A\) forces
\begin{equation}
  n=N_c .
  \label{eq:wzw-level-forced-by-matching}
\end{equation}
This is the same equality obtained above from the Wess--Zumino functional, now
read as a test of infrared phases.  If the chiral symmetry were unbroken and
the theory confined to a unique trivial gapped vacuum, there would be no
massless determinant line, no Goldstone WZW term, and no topological response;
that candidate would fail \eqref{eq:qcd-left-flavor-uv-anomaly-repeat}.  An
unbroken phase is not excluded by this argument if it contains massless
composite fermions or a topological/invertible sector carrying exactly the same
background cocycle.  The matching condition constrains the infrared
possibilities; it is not a proof of which dynamical phase QCD chooses.

The vector subgroup illustrates the same logic in a form that is often hidden
by notation.  Set \(L=R=V\) and couple a single vector background \(v\) to both
chiralities.  In the anomaly coefficient
\eqref{eq:flavor-anomaly-coefficient}, the left-handed and right-handed
Dirac components have the same vector generator.  Their contributions to the
vector current anomaly cancel:
\begin{equation}
  \mathcal A^{abc}_{V}
  =
  \frac{N_c}{2}
  \left[
    \operatorname{Tr}\bigl(T^a\{T^b,T^c\}\bigr)
    -
    \operatorname{Tr}\bigl(T^a\{T^b,T^c\}\bigr)
  \right]
  =
  0 .
  \label{eq:vector-flavor-anomaly-cancels}
\end{equation}
Thus \(SU(N_f)_V\) is an anomaly-free symmetry that can act on a purely
massive color-singlet sector without requiring additional massless fields or a
topological response.  By contrast, a purely left chiral subgroup has the
nonzero coefficient in \eqref{eq:qcd-left-flavor-uv-anomaly-repeat}; if it
remained unbroken in a gapped infrared theory, its anomaly would have to be
carried by some nontrivial infrared sector.

The positivity statement used earlier is logically separate but compatible
with this anomaly test.  Under Hypothesis~\ref{hyp:vafa-witten-measure-data},
equal positive quark masses select a vacuum preserving the vector flavor
group; taking the small-mass limit then aligns the chiral condensate with
\(SU(N_f)_V\) whenever the required limits exist.  Anomaly matching then
determines the response of the broken axial directions: the Goldstone target
\(SU(N_f)\) together with \(N_c\Gamma_{\mathrm{WZW}}\) reproduces the
ultraviolet chiral anomaly, while the unbroken vector subgroup remains
nonanomalous.  The two inputs have different status.  The vector-alignment
argument depends on Euclidean positivity and limiting assumptions; the anomaly
matching equation is an exact background-response constraint once the
ultraviolet symmetry and regulator convention are fixed.

For the electromagnetic specialization use the dual tensor defined at the start
of the chapter and the flavor-generator normalization just stated, and fix
\[
  T^3=\frac{\sigma_3}{2}
  =
  \frac12\operatorname{diag}(1,-1).
\]
Let \(q\) denote the dimensionless light-quark charge matrix, and embed
electromagnetism by
\[
  \ell_\mu=r_\mu=e\,q A_\mu,
  \qquad
  q=\operatorname{diag}(q_1,\ldots,q_{N_f}),
\]
The relevant microscopic diagram is the quark triangle in
Figure~\ref{fig:em-axial-flavor-triangle}.  The figure fixes the assignment of
the axial flavor generator to one vertex and the two electromagnetic current
insertions to the other vertices before the trace over light-quark charges is
evaluated.
\begin{figure}[H]
  \centering
  \begin{tikzpicture}[>=Stealth,scale=0.9,baseline={(0,-0.1)},
    photon/.style={decorate,decoration={snake,amplitude=1.4pt,segment length=6pt},
      line width=0.75pt}]
    \coordinate (v1) at (0,1.1);
    \coordinate (v2) at (-1.15,-0.85);
    \coordinate (v3) at (1.15,-0.85);
    \draw[thick] (v1)--(v2)--(v3)--cycle;
    \draw[->,thick] (-2.45,0.15)--(v2);
    \draw[photon,->] (v1)--(0,2.25);
    \draw[photon,->] (v3)--(2.25,-0.85);
    \fill (v1) circle (1.5pt);
    \fill (v2) circle (1.5pt);
    \fill (v3) circle (1.5pt);
    \node[left] at (-2.45,0.15) {\(j^{A\mu}_a\)};
    \node[above] at (0,2.25) {\(\gamma\)};
    \node[right] at (2.25,-0.85) {\(\gamma\)};
  \end{tikzpicture}
  \caption{The electromagnetic background probes the axial \(SU(2)\) flavor
  anomaly through the quark triangle with one axial-current insertion and two
  electromagnetic-current insertions.}
  \label{fig:em-axial-flavor-triangle}
\end{figure}
\begin{remark}[Electromagnetic specialization]
\label{rem:electromagnetic-anomaly-specialization}
The corresponding axial-flavor Ward identity is
\[
  \partial_\mu j_a^{A\mu}
  =
  -\frac{N_c e^2}{16\pi^2}
  \operatorname{Tr}\!\left(T^a\{q,q\}\right)
  F_{\mu\nu}\widetilde F^{\mu\nu}.
\]
The anticommutator records the two identical electromagnetic insertions in the
triangle; since the charge matrix is diagonal,
\[
  \operatorname{Tr}\!\left(T^a\{q,q\}\right)
  =
  2\operatorname{Tr}\!\left(T^a q^2\right).
\]
Equivalently,
\[
  \partial_\mu j_a^{A\mu}
  =
  -\frac{N_c e^2}{8\pi^2}
  \operatorname{Tr}\!\left(T^a q^2\right)
  F_{\mu\nu}\widetilde F^{\mu\nu}.
\]
This is the same anomaly coefficient as the preceding chapter's
Dirac-fermion formula, rewritten from
\(\epsilon^{\mu\nu\rho\sigma}F_{\mu\nu}F_{\rho\sigma}\) to
\(2F_{\mu\nu}\widetilde F^{\mu\nu}\).
For \(q=\operatorname{diag}(2/3,-1/3)\), the trace is nonzero only for
\(a=3\), where
\[
  \operatorname{Tr}\!\left(T^3 q^2\right)=\frac16.
\]
Thus
\[
  \operatorname{Tr}\!\left(T^3\{q,q\}\right)=\frac13 .
\]
Consequently
the gauged WZW functional contains the neutral-pion coupling
\begin{equation}
  \mathcal L_{\pi^0\gamma\gamma}
  =
  \frac{N_c e^2}{16\pi^2 f_\pi}
  \operatorname{Tr}\!\left(T^3\{q,q\}\right)\,
  \pi^0 F_{\mu\nu}\widetilde F^{\mu\nu},
  \label{eq:pi-zero-two-photon-anomaly}
\end{equation}
namely
\[
  \mathcal L_{\pi^0\gamma\gamma}
  =
  \frac{N_c e^2}{48\pi^2 f_\pi}
  \pi^0 F_{\mu\nu}\widetilde F^{\mu\nu}.
\]
For \(N_c=3\), the coefficient multiplying
\(\pi^0F_{\mu\nu}\widetilde F^{\mu\nu}\) is \(e^2/(16\pi^2f_\pi)\).
The coefficient is the low-energy coefficient required by matching the
microscopic electromagnetic flavor anomaly.
\end{remark}

Figure~\ref{fig:pi-zero-two-photon-decay} records the low-energy vertex matched
from that microscopic anomaly: one neutral pion couples to two photons with the
coefficient fixed above.

\begin{figure}[H]
  \centering
  \begin{tikzpicture}[>=Stealth,scale=0.9,baseline={(0,-0.1)},
    photon/.style={decorate,decoration={snake,amplitude=1.4pt,segment length=6pt},
      line width=0.75pt}]
    \draw[thick,dashed,->] (-2.0,0) -- (0,0);
    \draw[photon,->] (0,0) -- (1.75,0.75);
    \draw[photon,->] (0,0) -- (1.75,-0.75);
    \fill (0,0) circle (2pt);
    \node[left] at (-2.0,0) {\(\pi^0\)};
    \node[right] at (1.75,0.75) {\(\gamma\)};
    \node[right] at (1.75,-0.75) {\(\gamma\)};
  \end{tikzpicture}
  \caption{The electromagnetic flavor anomaly produces the local
  \(\pi^0\gamma\gamma\) vertex in the pion effective action.}
  \label{fig:pi-zero-two-photon-decay}
\end{figure}

Writing
\[
  \mathcal L_{\pi^0\gamma\gamma}
  =
  \frac{g_{\pi\gamma\gamma}}4\,
  \pi^0 F_{\mu\nu}\widetilde F^{\mu\nu}
\]
gives \(g_{\pi\gamma\gamma}=e^2/(4\pi^2f_\pi)=\alpha/(\pi f_\pi)\) for
\(N_c=3\).  The corresponding decay rate is
\[
  \Gamma(\pi^0\to2\gamma)
  =
  \frac{\alpha^2m_{\pi^0}^3}{64\pi^3f_\pi^2},
\]
in the convention for \(f_\pi\) used above.

\section{Conceptual Output of the Gauge-Dynamics Sequence}

The dependencies established in this sequence are as follows.  Gauge fields are
local connections.  Gauge fixing supplies local coordinates and ghosts.  BRST
cohomology defines physical states, local operators, counterterms, and anomaly
candidates in the gauge-fixed formalism.  QCD renormalization gives
asymptotically free short-distance variables.  Local anomalies are constrained
by Wess--Zumino consistency and descent.  Finite background anomalies are
encoded by cocycles on background-field space.  Exact anomaly classes are
preserved along RG flow.  Spontaneous symmetry breaking realizes these
constraints through Goldstone fields and the Wess--Zumino--Witten functional.
