\chapter{Subdivergences and Forest Formulas}
\label{ch:volume-ii-subdivergences-forest-formulas}
\ConventionsEuclidean

\section{Path-Integral Status: Formal Coefficientwise Renormalization}

The construction in this chapter is a perturbative construction extracted from
a regulated Euclidean source functional.  Let \(\langle-\rangle_{0,\Lambda}\)
be a specified ultraviolet-regulated Gaussian expectation and let
\(V_\Lambda\) be a finite linear combination of regulated local interaction
functionals with couplings
\[
  g=(g_1,\ldots,g_r),
  \qquad
  V_\Lambda[\phi]=\sum_{\alpha=1}^r g_\alpha O_{\alpha,\Lambda}[\phi],
\]
where the subscript \(\Lambda\) denotes the ultraviolet regulator used to
define both the covariance and the inserted local functionals.  The normalized
source functional is
\[
  Z_{\Lambda,g}[J]
  =
  \frac{\left\langle
    \exp\big(\langle J,\phi\rangle -V_\Lambda[\phi]\big)
  \right\rangle_{0,\Lambda}}
  {\left\langle
    \exp\big(-V_\Lambda[\phi]\big)
  \right\rangle_{0,\Lambda}},
\]
understood as an identity in the formal power-series ring
\(\mathbb C[[g_1,\ldots,g_r]]\).  Its correlation functions are expanded as
formal power series in \(g\):
\[
  \big\langle \phi(x_1)\cdots\phi(x_n)\big\rangle_{\Lambda,g}
  =
  \sum_{\nu\in\mathbb N^r}
  g^\nu\,
  G^{(\nu)}_{\Lambda,n}(x_1,\ldots,x_n),
  \qquad
  g^\nu=\prod_{\alpha=1}^r g_\alpha^{\nu_\alpha}.
\]
Each coefficient \(G^{(\nu)}_{\Lambda,n}\) is a finite sum of graph
distributions, or equivalently of momentum-space graph integrals after a
choice of momentum routing and test functions.  BPHZ renormalization acts on
these coefficient distributions graph by graph.  For a graph \(\Gamma\), the
forest formula constructs a renormalized coefficient \(R_\Gamma I_\Gamma\) by
subtracting Taylor polynomials in the external momenta of divergent subgraphs.
Only after this coefficientwise operation has been performed does one assemble
the formal series again.

Thus BPHZ, as used here, is not an assertion that the formal symbol
\(\mathcal D\phi\) denotes a countably additive measure on an unregulated
continuum field space, and it is not a nonperturbative construction of the
theory.  It is a precise local subtraction scheme for the formal perturbative
expansion of regulated Euclidean Green functions and 1PI kernels.  Its
comparison with Wilsonian effective actions is therefore an order-by-order
comparison between two coordinate systems on perturbative coefficient data,
not an equality of nonperturbative path integrals.


\section{Graphs and Ultraviolet Subgraphs}

Work in a massive Euclidean scalar theory in \(D\) dimensions, with local
polynomial interactions and a fixed choice of momentum routing.  The massive
assumption isolates ultraviolet questions from infrared ones; the subtraction
points will also be chosen away from exceptional momentum configurations.  The
graphs in this chapter represent 1PI kernels; connected Green functions are
reconstructed from 1PI kernels by the Legendre tree expansion discussed
earlier.  A connected one-particle-irreducible Feynman graph \(\Gamma\) has a
finite set \(E(\Gamma)\) of internal lines, a finite set \(V(\Gamma)\) of
vertices, and external momenta
\[
  p=(p_1,\ldots,p_N),
  \qquad
  \sum_{a=1}^N p_a=0 .
\]
Choose independent loop momenta
\(\ell=(\ell_1,\ldots,\ell_{L(\Gamma)})\), where
\[
  L(\Gamma)=|E(\Gamma)|-|V(\Gamma)|+1 .
\]
With polynomial numerator \(N_\Gamma(\ell,p)\) and propagator masses
\(m_e>0\), the Euclidean momentum-space integrand has the form
\[
  I_\Gamma(\ell,p)
  =
  \frac{N_\Gamma(\ell,p)}
       {\prod_{e\in E(\Gamma)}
        \big(q_e(\ell,p)^2+m_e^2\big)} .
\]
Here \(q_e(\ell,p)\) is the momentum carried by the line \(e\), determined by
the chosen momentum routing.  Coupling constants, symmetry factors, and
overall powers of \(2\pi\) are suppressed in this chapter because they do not
affect the ultraviolet degree or locality of the subtraction.

\begin{definition}[Momentum subgraph]\leavevmode\par
A momentum subgraph \(\gamma\subseteq\Gamma\) is specified by a nonempty subset
\(E(\gamma)\subseteq E(\Gamma)\), together with all vertices incident to those
internal lines.  Its connected components are formed with the incidence
relation inherited from \(\Gamma\).  The loop number of \(\gamma\) is
\[
  L(\gamma)=|E(\gamma)|-|V(\gamma)|+c(\gamma),
\]
where \(c(\gamma)\) is the number of connected components of \(\gamma\).
The external half-lines of \(\gamma\) are all half-lines incident to
\(V(\gamma)\) that are not internal lines of \(\gamma\).  They include original
external half-lines of \(\Gamma\) and the half-lines obtained by cutting
internal lines of \(\Gamma\setminus\gamma\) at vertices of \(\gamma\).  The
momenta on these half-lines are denoted \(p_\gamma\), with momentum
conservation imposed separately on each connected component.  The contracted
graph \(\Gamma/\gamma\) is obtained by replacing every connected component of
\(\gamma\) by a single vertex carrying the total external momentum of that
component.
\end{definition}

The ultraviolet scaling associated with \(\gamma\) scales independent loop
momenta internal to \(\gamma\) by a common parameter \(\rho\) and keeps
\(p_\gamma\) fixed.  The momenta \(p_\gamma\) may include loop momenta of the
contracted graph; during the test of \(\gamma\) they are treated as external
parameters.  If the leading homogeneous scaling of \(I_\gamma\) together with
the internal loop measure is
\[
  \dd^{DL(\gamma)}\ell_\gamma\,
  I_\gamma(\rho\ell_\gamma,p_\gamma)
  =
  O(\rho^{\omega(\gamma)})
  \dd^{DL(\gamma)}\ell_\gamma
  \qquad
  (\rho\to\infty),
\]
then \(\omega(\gamma)\) is the superficial degree of divergence.  For a scalar
graph with numerator degree \(r(\gamma)\),
\[
  \omega(\gamma)
  =
  D L(\gamma)+r(\gamma)-2|E(\gamma)|.
\]
Equivalently, \(\omega(\gamma)\) is the mass dimension of the local polynomial
that can be produced when \(\gamma\) is shrunk to a point.  A connected
subgraph is one-particle-irreducible if it remains connected after any one of
its internal lines is removed.  A proper connected one-particle-irreducible
subgraph with \(\omega(\gamma)\ge0\) is called a divergent subgraph; proper
means \(\gamma\ne\Gamma\).  A disconnected subgraph is divergent when each of
its connected components is divergent.  In that case its local counterterm is
the product of the counterterms of its connected components.

\section{Taylor Operators and Local Vertices}

Fix for each divergent subgraph \(\gamma\) a subtraction point
\(p_\gamma^\ast\) in the space of independent external momenta of \(\gamma\).
The point is chosen away from infrared singularities; for massive Euclidean
graphs, zero momentum is allowed when the relevant Green function is infrared
regular.  Let \(a=(a_1,\ldots,a_M)\) be affine coordinates for the independent
components of \(p_\gamma-p_\gamma^\ast\), after imposing momentum
conservation.  The Taylor operator of degree \(\omega(\gamma)\) is
\[
  (t_\gamma f)(p_\gamma)
  =
  \sum_{|\alpha|\le\omega(\gamma)}
  \frac{a^\alpha}{\alpha!}
  \left.
    \partial_a^\alpha f(p_\gamma^\ast+a)
  \right|_{a=0}.
\]
It acts on the dependence of the integrand on the external momenta of
\(\gamma\), while all loop momenta internal to \(\gamma\) remain integration
variables and all momenta belonging to the contracted graph are held fixed.
For a disconnected divergent subgraph, \(t_\gamma\) is the product of the
Taylor operators for its connected components.

The image \(t_\gamma I_\Gamma\) is polynomial in \(p_\gamma\) of degree at most
\(\omega(\gamma)\), with coefficients depending on the loop variables internal
to \(\gamma\) and on the remaining variables of the contracted graph.  After
the internal variables of \(\gamma\) have been integrated or recursively
prepared, this polynomial becomes a vertex factor in \(\Gamma/\gamma\).  In
position space, a polynomial in the momentum entering a contracted vertex is a
finite linear combination of derivatives acting at that vertex.  Therefore
\(t_\gamma I_\Gamma\) represents a local insertion in \(\Gamma/\gamma\).  This
is the graph-level form of the locality statement established in the preceding
chapter.

Figure~\ref{fig:volume-ii-taylor-subgraph-local-vertex} shows this graph-level
operation explicitly: Taylor expansion in the external variables of
\(\gamma\) becomes a polynomial counterterm vertex in the contracted graph
\(\Gamma/\gamma\).

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[>=Stealth,x=1cm,y=1cm,every node/.style={font=\small}]
    \tikzset{
      line/.style={line width=0.95pt},
      arr/.style={-{Latex[length=2mm]},line width=0.9pt},
      blob/.style={ellipse,draw=black,fill=gray!15,line width=0.9pt,
        minimum width=1.35cm,minimum height=0.72cm}
    }
    \begin{scope}
      \node[font=\normalsize] at (0,2.0) {subgraph};
      \draw[line] (-2.45,0.55)--(-1.2,0.55);
      \node[blob] (G) at (0,0.55) {\(\gamma\)};
      \draw[line] (G.east)--(2.45,0.55);
      \draw[line] (-0.35,0.05)--(-0.85,-0.95);
      \draw[line] (0.35,0.05)--(0.85,-0.95);
      \node[blue!70!black,align=center] at (0,-1.55)
        {external variables \(p_\gamma\)};
    \end{scope}

    \draw[arr] (3.0,0.55)--(4.35,0.55);
    \node[above] at (3.68,0.72) {\(t_\gamma\)};

    \begin{scope}[shift={(6.3,0)}]
      \node[font=\normalsize] at (0,2.0) {local vertex};
      \draw[line] (-2.45,0.55)--(-0.22,0.55);
      \draw[line] (0.22,0.55)--(2.45,0.55);
      \fill[red!75!black] (0,0.55) circle (3pt);
      \draw[line] (-0.05,0.2)--(-0.75,-0.95);
      \draw[line] (0.05,0.2)--(0.75,-0.95);
      \node[red!75!black,align=center] at (0,-1.55)
        {polynomial in \(p_\gamma\)\\degree \(\le\omega(\gamma)\)};
    \end{scope}

    \draw[arr] (9.15,0.55)--(10.5,0.55);

    \begin{scope}[shift={(12.7,0)}]
      \node[font=\normalsize] at (0,2.0) {contracted graph};
      \draw[line] (-2.1,0.55)--(-0.3,0.55);
      \draw[line] (0.3,0.55)--(2.1,0.55);
      \draw[line] (-0.2,0.18)--(-1.05,-0.95);
      \draw[line] (0.2,0.18)--(1.05,-0.95);
      \node[circle,draw=black,fill=red!15,line width=0.9pt,
        minimum size=0.55cm] at (0,0.55) {\(C_\gamma\)};
      \node[green!45!black,align=center] at (0,-1.55)
        {counterterm insertion\\in \(\Gamma/\gamma\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The Taylor operator \(t_\gamma\) extracts the polynomial part of a
  divergent subgraph in its external variables.  The polynomial is represented
  in the contracted graph by a local counterterm vertex.}
  \label{fig:volume-ii-taylor-subgraph-local-vertex}
\end{figure}

\section{Spinneys, Forests, and the Recursive R-Operation}

\begin{definition}[Spinneys and forests]
A spinney \(S\) of \(\Gamma\) is a finite set of mutually disjoint connected
divergent proper subgraphs of \(\Gamma\).  Mutually disjoint means that no
internal line or vertex is shared.  The empty set is allowed.  A forest \(F\)
of \(\Gamma\) is a finite set of connected divergent subgraphs such that any
two elements of \(F\) are either disjoint or nested:
\[
  \gamma_1\cap\gamma_2=\emptyset,
  \qquad
  \gamma_1\subset\gamma_2,
  \qquad
  \hbox{or}
  \qquad
  \gamma_2\subset\gamma_1 .
\]
The full graph \(\Gamma\) is included among the possible forest elements when
\(\Gamma\) is one-particle-irreducible and \(\omega(\Gamma)\ge0\).  The empty
forest is allowed.
\end{definition}

Figure~\ref{fig:volume-ii-compatible-forests} fixes the compatibility relation
used in the forest sum: divergent subgraphs may be disjoint or nested, whereas
overlapping subgraphs appear in separate forest terms.

\begin{figure}[H]
  \centering
  \resizebox{0.94\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\small}]
    \tikzset{line/.style={line width=0.95pt}}
    \begin{scope}
      \node[font=\normalsize] at (0,2.35) {disjoint};
      \draw[line] (-2.4,0.55)--(2.4,0.55);
      \draw[line] (-1.8,0.55)..controls (-1.3,1.2) and (-0.7,1.2)..(-0.2,0.55);
      \draw[line] (-1.8,0.55)..controls (-1.3,-0.1) and (-0.7,-0.1)..(-0.2,0.55);
      \draw[line] (0.2,0.55)..controls (0.7,1.2) and (1.3,1.2)..(1.8,0.55);
      \draw[line] (0.2,0.55)..controls (0.7,-0.1) and (1.3,-0.1)..(1.8,0.55);
      \draw[blue!70!black,dashed,line width=1pt] (-1.95,-0.35) rectangle (-0.05,1.28);
      \draw[green!50!black,dashed,line width=1pt] (0.05,-0.35) rectangle (1.95,1.28);
      \node[blue!70!black] at (-1.0,-0.8) {\(\gamma_1\)};
      \node[green!50!black] at (1.0,-0.8) {\(\gamma_2\)};
    \end{scope}

    \begin{scope}[shift={(5.2,0)}]
      \node[font=\normalsize] at (0,2.35) {nested};
      \draw[line] (-2.4,0.55)--(2.4,0.55);
      \draw[line] (-1.8,0.55)..controls (-0.9,1.45) and (0.9,1.45)..(1.8,0.55);
      \draw[line] (-1.8,0.55)..controls (-0.9,-0.35) and (0.9,-0.35)..(1.8,0.55);
      \draw[line] (-0.65,0.55)..controls (-0.25,0.98) and (0.25,0.98)..(0.65,0.55);
      \draw[line] (-0.65,0.55)..controls (-0.25,0.12) and (0.25,0.12)..(0.65,0.55);
      \draw[blue!70!black,dashed,line width=1pt] (-2.0,-0.55) rectangle (2.0,1.55);
      \draw[green!50!black,dashed,line width=1pt] (-0.82,0.0) rectangle (0.82,1.08);
      \node[blue!70!black] at (-1.65,-0.95) {\(\gamma_2\)};
      \node[green!50!black] at (0,-0.38) {\(\gamma_1\)};
    \end{scope}

    \begin{scope}[shift={(10.4,0)}]
      \node[font=\normalsize] at (0,2.35) {overlapping};
      \draw[line] (-2.4,0.55)--(2.4,0.55);
      \draw[line] (-1.8,0.55)..controls (-1.0,1.35) and (-0.1,1.35)..(0.35,0.55);
      \draw[line] (-1.8,0.55)..controls (-1.0,-0.25) and (-0.1,-0.25)..(0.35,0.55);
      \draw[line] (-0.35,0.55)..controls (0.1,1.35) and (1.0,1.35)..(1.8,0.55);
      \draw[line] (-0.35,0.55)..controls (0.1,-0.25) and (1.0,-0.25)..(1.8,0.55);
      \draw[blue!70!black,dashed,line width=1pt] (-1.95,-0.45) rectangle (0.55,1.42);
      \draw[green!50!black,dashed,line width=1pt] (-0.55,-0.45) rectangle (1.95,1.42);
      \node[blue!70!black] at (-1.05,-0.88) {\(\gamma_1\)};
      \node[green!50!black] at (1.05,-0.88) {\(\gamma_2\)};
      \node[red!70!black,align=center] at (0,-1.55)
        {separate forest terms};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{A forest contains compatible divergent subgraphs: disjoint or
  nested collections are compatible.  Overlapping subgraphs are handled by
  separate forest terms.}
  \label{fig:volume-ii-compatible-forests}
\end{figure}

The recursive construction is most transparent when written in two steps.
First define the prepared integrand
\[
  \overline R_\Gamma I_\Gamma
  =
  I_\Gamma
  +
  \sum_{\substack{S\ {\rm spinney\ of}\ \Gamma\\S\ne\emptyset}}
  \left(\prod_{\gamma\in S} C_\gamma\right) I_\Gamma .
\]
The product of \(C_\gamma\)'s means: for each disjoint component
\(\gamma\in S\), replace the prepared subgraph by its local counterterm vertex
and insert the resulting product of local vertices into the contracted graph
\(\Gamma/S\).  Since the elements of a spinney are disjoint, these
replacements can be made simultaneously.  The counterterm attached to a
connected divergent graph is defined recursively by
\[
  C_\gamma I_\gamma
  =
  -\,t_\gamma\,\overline R_\gamma I_\gamma .
\]
Thus the counterterm of \(\gamma\) is the negative Taylor polynomial of the
subgraph after its own proper subdivergences have already been prepared.  The
recursion starts at divergent graphs with no proper divergent subgraphs, for
which \(\overline R_\gamma I_\gamma=I_\gamma\).  For disconnected divergent
subgraphs, \(C_\gamma\) is the product over connected components.

The full \(R\)-operation is
\[
  R_\Gamma I_\Gamma
  =
  \begin{cases}
    (1-t_\Gamma)\,\overline R_\Gamma I_\Gamma,
      & \omega(\Gamma)\ge0,\\[4pt]
    \overline R_\Gamma I_\Gamma,
      & \omega(\Gamma)<0.
  \end{cases}
\]
This formula enforces the order of operations: proper subdivergences are
subtracted inside \(\overline R_\Gamma\), and the overall Taylor subtraction
is applied afterward when the whole graph is superficially divergent.

\begin{proposition}[Bogoliubov preparation recursion]
\label{prop:bogoliubov-preparation-recursion}
Fix the subtraction points and Taylor degrees described above.  For every
connected 1PI graph \(\Gamma\), the recursive definitions of
\(\overline R_\Gamma\), \(C_\gamma\), and \(R_\Gamma\) determine a finite
linear combination of integrands obtained from \(\Gamma\) by replacing
divergent connected subgraphs by local polynomial vertices.  If
\(\gamma\) is connected and divergent, then \(C_\gamma I_\gamma\) is a
polynomial in the external momenta \(p_\gamma\) of degree at most
\(\omega(\gamma)\), with coefficients that are finite linear combinations of
prepared lower-loop or lower-line graph integrands.
\end{proposition}

\begin{proof}
Order connected graphs by the number of internal lines.  A proper connected
subgraph of \(\Gamma\) has strictly fewer internal lines than \(\Gamma\), so
the recursion is well founded.  For graphs with no proper divergent connected
subgraphs, \(\overline R_\Gamma I_\Gamma=I_\Gamma\), and, if \(\Gamma\) is
divergent, \(C_\Gamma I_\Gamma=-t_\Gamma I_\Gamma\), which is a polynomial in
the external variables by the definition of \(t_\Gamma\).

Assume the claim for all proper connected divergent subgraphs of
\(\Gamma\).  A spinney is a finite set of mutually disjoint such subgraphs,
so the product \(\prod_{\gamma\in S}C_\gamma\) can be inserted
simultaneously: each factor has already been constructed by the induction
hypothesis, and disjoint replacements affect disjoint vertex and line sets.
The number of spinneys is finite because \(\Gamma\) has finitely many internal
lines.  Hence \(\overline R_\Gamma I_\Gamma\) is a finite linear combination
of integrands on contracted graphs with local polynomial vertices.

If \(\Gamma\) itself is divergent, \(C_\Gamma I_\Gamma\) is defined as
\(-t_\Gamma\overline R_\Gamma I_\Gamma\).  The Taylor operator acts only on
the external variables of \(\Gamma\) and truncates at degree
\(\omega(\Gamma)\); local vertices produced inside \(\overline R_\Gamma\)
are polynomial factors in slower external variables and remain local after
the Taylor extraction.  Thus the new counterterm is again a local polynomial
vertex of the stated degree.  This completes the induction.
\end{proof}

\section{The Forest Formula}

Let \(\mathfrak F(\Gamma)\) be the set of forests whose elements are connected
divergent subgraphs of \(\Gamma\), including \(\Gamma\) itself when
\(\Gamma\) is one-particle-irreducible and \(\omega(\Gamma)\ge0\).  For
\(F\in\mathfrak F(\Gamma)\), define \(t_F\) by applying \(t_\gamma\) from
smaller subgraphs to larger subgraphs, with each smaller subtraction regarded
as a local vertex before the larger Taylor operator is applied.  If two
elements are disjoint, their operators commute.  The forest formula is the
following identity.
\begin{theorem}[Zimmermann forest formula]
\label{thm:zimmermann-forest-formula}
For the BPHZ recursion of
Proposition~\ref{prop:bogoliubov-preparation-recursion},
\begin{equation*}
  R_\Gamma I_\Gamma
  =
  \sum_{F\in\mathfrak F(\Gamma)}
  (-1)^{|F|}\,t_F I_\Gamma ,
\tag{\(\ast\)}
\label{eq:zimmermann-forest-formula}
\end{equation*}
with \(t_\emptyset=1\).  The identity is an identity of formal graph
integrands before loop integrations are performed.  In a nested chain,
Taylor operators are ordered from smaller subgraphs to larger subgraphs; for
disjoint subgraphs the order is immaterial.
\end{theorem}

\begin{proof}
We prove the identity by induction on the number of internal lines.  For a
graph with no proper divergent connected subgraphs, the spinney sum contains
only the empty spinney.  If \(\Gamma\) is superficially convergent then
\(R_\Gamma I_\Gamma=I_\Gamma\), which is the forest formula with only the
empty forest.  If \(\Gamma\) is superficially divergent then
\[
  R_\Gamma I_\Gamma=(1-t_\Gamma)I_\Gamma
  =
  I_\Gamma-t_\Gamma I_\Gamma ,
\]
which is the sum over the two forests
\(\emptyset\) and \(\{\Gamma\}\).

Assume the statement for every proper connected divergent subgraph of
\(\Gamma\).  Expanding the counterterm of a proper connected divergent
subgraph \(\gamma\) gives
\begin{equation}
  C_\gamma I_\gamma
  =
  -t_\gamma \overline R_\gamma I_\gamma
  =
  \sum_{\substack{F_\gamma\in\mathfrak F(\gamma)\\
                  \gamma\in F_\gamma}}
  (-1)^{|F_\gamma|}t_{F_\gamma}I_\gamma .
  \label{eq:bphz-counterterm-as-forests-ending-at-gamma}
\end{equation}
Indeed, by the induction hypothesis, \(\overline R_\gamma\) is the forest sum
over forests of \(\gamma\) not containing \(\gamma\); multiplying by
\(-t_\gamma\) appends \(\gamma\) as the largest element and changes the sign.

Now expand the spinney sum in \(\overline R_\Gamma\).  A nonempty spinney
\(S\) is a set of pairwise disjoint proper connected divergent subgraphs.
For each \(\gamma\in S\), choose a forest \(F_\gamma\) of \(\gamma\) whose
largest element is \(\gamma\), as in
\eqref{eq:bphz-counterterm-as-forests-ending-at-gamma}.  The union
\[
  F=\bigcup_{\gamma\in S}F_\gamma
\]
is a forest of \(\Gamma\) that does not contain \(\Gamma\), because different
members of \(S\) are disjoint and all further elements lie inside those
members.  Conversely, every forest of \(\Gamma\) not containing \(\Gamma\)
has a unique set of maximal elements; these maximal elements form a spinney,
and the portions lying below each maximal element are precisely the
corresponding \(F_\gamma\)'s.  The product of the signs in the spinney
expansion is
\((-1)^{\sum_{\gamma\in S}|F_\gamma|}=(-1)^{|F|}\).
Therefore
\begin{equation}
  \overline R_\Gamma I_\Gamma
  =
  \sum_{\substack{F\in\mathfrak F(\Gamma)\\\Gamma\notin F}}
  (-1)^{|F|}t_F I_\Gamma .
  \label{eq:bphz-prepared-as-forests-not-containing-gamma}
\end{equation}

If \(\Gamma\) is superficially convergent, \(R_\Gamma=\overline R_\Gamma\)
and this is the desired formula.  If \(\Gamma\) is superficially divergent,
then
\[
  R_\Gamma I_\Gamma
  =
  (1-t_\Gamma)\overline R_\Gamma I_\Gamma .
\]
The first term is
\eqref{eq:bphz-prepared-as-forests-not-containing-gamma}.  The second term
appends \(\Gamma\) as the unique largest element to every forest not already
containing \(\Gamma\), with one extra minus sign.  This gives exactly the
remaining forests, namely those containing \(\Gamma\), and proves
\eqref{eq:zimmermann-forest-formula}.
\end{proof}

The inclusion-exclusion structure can be seen directly in a graph with one
proper divergent subgraph \(\gamma\) and a divergent overall graph \(\Gamma\).
The possible forests are
\[
  \emptyset,\qquad
  \{\gamma\},\qquad
  \{\Gamma\},\qquad
  \{\gamma,\Gamma\}.
\]
The corresponding operator is
\[
  R_\Gamma
  =
  1-t_\gamma-t_\Gamma+t_\Gamma t_\gamma
  =
  (1-t_\Gamma)(1-t_\gamma),
\]
where \(t_\gamma\) acts before \(t_\Gamma\).  The last term restores the part
counted twice at the boundary where the subgraph and the whole graph shrink in
a nested way.

\section{Order of Subtractions and Scale Hierarchies}

The prescription that \(t_\gamma\) acts before \(t_\Gamma\) for
\(\gamma\subset\Gamma\) is determined by the geometry of ultraviolet limits.  A
nested forest represents a hierarchy of large internal momenta.  For example,
if loop momenta internal to \(\gamma\) are of size \(\rho\sigma\) and the
remaining loop momenta of \(\Gamma/\gamma\) are of size \(\rho\), with
\(\sigma\to\infty\) before \(\rho\to\infty\), then the innermost singular
region is the subgraph \(\gamma\).  Its Taylor polynomial is first converted
into a local vertex of the contracted graph.  After that replacement, the
overall Taylor operator acts on the local integrand of the larger graph.

This ordering can be expressed algebraically.  If
\[
  C_\gamma I_\gamma=-t_\gamma\overline R_\gamma I_\gamma,
\]
then the term \(t_\Gamma t_\gamma I_\Gamma\) in a forest is interpreted as
\[
  t_\Gamma\bigl((t_\gamma I_\gamma)\,I_{\Gamma/\gamma}\bigr),
\]
where \(t_\gamma I_\gamma\) is a polynomial in the external momenta entering
\(\gamma\).  The polynomial is part of the local vertex data of
\(\Gamma/\gamma\), so the overall subtraction is again a Taylor polynomial in
the external momenta of \(\Gamma\).  This is the mechanism by which nested
subtractions preserve locality.

Overlapping divergent subgraphs are not disjoint and neither contains the
other, so they are not placed in the same forest.  Their common ultraviolet
region is represented by the sum over compatible forests together with the
overall subtraction.  The forest sum is the inclusion-exclusion expression that
subtracts all ultraviolet boundary regions while assigning counterterms only
to local contracted subgraphs.

\section{The Two-Loop Self-Energy Forest}

Return to the two-loop self-energy graph in the six-dimensional cubic theory.
Let \(\Gamma\) be the full graph and let \(\gamma_L,\gamma_R\) denote the left
and right one-loop self-energy subgraphs.  These two subgraphs overlap: they
share the central propagator.  Therefore the forest set is
\[
  \emptyset,\quad
  \{\gamma_L\},\quad
  \{\gamma_R\},\quad
  \{\Gamma\},\quad
  \{\gamma_L,\Gamma\},\quad
  \{\gamma_R,\Gamma\}.
\]
The overlap of \(\gamma_L\) and \(\gamma_R\) is represented by the two separate
one-subgraph forests \(\{\gamma_L\}\) and \(\{\gamma_R\}\), together with their
nested overall forests.

For this graph the \(R\)-operation is
\[
\begin{aligned}
  R_\Gamma I_\Gamma
  &=
  I_\Gamma
  -t_{\gamma_L}I_\Gamma
  -t_{\gamma_R}I_\Gamma
  -t_\Gamma I_\Gamma        \\
  &\quad
  +t_\Gamma t_{\gamma_L}I_\Gamma
  +t_\Gamma t_{\gamma_R}I_\Gamma .
\end{aligned}
\]
The first two subgraph subtraction terms remove the two Schwinger-parameter
boundary regions
\[
  \alpha_1,\alpha_2,\alpha_5\to0,
  \qquad
  \alpha_3,\alpha_4,\alpha_5\to0 .
\]
After these subtractions, the remaining overall boundary
\(\alpha_1,\ldots,\alpha_5\to0\) is subtracted by \(t_\Gamma\).  Since
\(\omega(\Gamma)=2\) for the self-energy in \(D=6\), \(t_\Gamma\) extracts the
constant and quadratic terms in the external momentum.  The final counterterm
is therefore a linear combination of the local operators \(\phi^2\) and
\(\partial_\mu\phi\,\partial_\mu\phi\).

Figure~\ref{fig:volume-ii-diamond-forest-formula} displays the diamond graph in
which the two one-loop subdivergences overlap; the forest formula subtracts
each compatible option and then the remaining overall divergence.

\begin{figure}[H]
  \centering
  \resizebox{0.97\textwidth}{!}{%
  \begin{tikzpicture}[>=Stealth,x=1cm,y=1cm,every node/.style={font=\small}]
    \tikzset{line/.style={line width=0.95pt}}
    \begin{scope}
      \node[font=\normalsize] at (0,2.5) {overlapping subdivergences};
      \draw[line] (-3.0,0.65)--(-2.1,0.65);
      \draw[line] (-2.1,0.65)..controls (-1.35,1.55) and (-0.45,1.55)..(0.0,0.65);
      \draw[line] (-2.1,0.65)..controls (-1.35,-0.25) and (-0.45,-0.25)..(0.0,0.65);
      \draw[line] (0.0,0.65)..controls (0.45,1.55) and (1.35,1.55)..(2.1,0.65);
      \draw[line] (0.0,0.65)..controls (0.45,-0.25) and (1.35,-0.25)..(2.1,0.65);
      \draw[line] (2.1,0.65)--(3.0,0.65);
      \draw[blue!70!black,dashed,line width=1pt] (-2.32,-0.47) rectangle (0.28,1.72);
      \draw[green!50!black,dashed,line width=1pt] (-0.28,-0.47) rectangle (2.32,1.72);
      \draw[red!70!black,dotted,line width=1.2pt] (-2.62,-0.78) rectangle (2.62,2.0);
      \node[blue!70!black] at (-1.15,-0.9) {\(\gamma_L\)};
      \node[green!50!black] at (1.15,-0.9) {\(\gamma_R\)};
      \node[red!70!black] at (0,2.18) {\(\Gamma\)};
    \end{scope}

    \begin{scope}[shift={(7.5,0)}]
      \node[font=\normalsize] at (0,2.5) {forest terms};
      \node[align=left,anchor=west] at (-2.6,1.45)
        {\(\emptyset\)};
      \node[align=left,anchor=west] at (-2.6,0.95)
        {\(\{\gamma_L\},\ \{\gamma_R\}\)};
      \node[align=left,anchor=west] at (-2.6,0.45)
        {\(\{\Gamma\}\)};
      \node[align=left,anchor=west] at (-2.6,-0.05)
        {\(\{\gamma_L,\Gamma\},\ \{\gamma_R,\Gamma\}\)};
      \draw[red!70!black,line width=0.9pt] (-2.55,-0.58)--(2.35,-0.58);
      \node[red!70!black,align=center] at (0,-1.08)
        {overlap represented separately};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The two one-loop subdivergences of the diamond graph overlap along
  one internal line.  The forest formula treats them in separate compatible
  forests, while the overall graph is subtracted after either one-loop
  subgraph subtraction.}
  \label{fig:volume-ii-diamond-forest-formula}
\end{figure}

\section{Finiteness and Locality}

\begin{lemma}[Hepp-sector forest associated to an ultraviolet hierarchy]
\label{lem:hepp-sector-forest-extraction}
Fix a smooth Hepp-sector partition of the noncompact loop-momentum region of
a graph \(\Gamma\), so that in each sector the magnitudes of the line momenta
are ordered up to bounded ratios.  This is a sector-local combinatorial
statement: it assumes that the analytic sector charts, adapted loop bases, and
uniform comparison constants have already been supplied.  In one sector, the
ultraviolet boundary strata determine a finite forest \(F_S\) as follows: for
each threshold in the ordered list of large line momenta, form the subgraph
spanned by lines above the threshold, decompose it into connected components,
keep the one-particle-irreducible components that contain loops, and then
discard components with negative superficial degree.  The retained components
are pairwise disjoint or nested.  Moreover, every superficially divergent
ultraviolet boundary of that sector is represented by a member of this forest,
or by a quotient graph obtained after members of the forest have been
contracted.
\end{lemma}

\begin{proof}
Label the distinct momentum-size thresholds in the sector by
\[
  \rho_1\gg \rho_2\gg\cdots\gg \rho_N
\]
after subdividing the partition of unity if necessary so that the ratios
\(\rho_{j+1}/\rho_j\) are the sector coordinates.  Let \(H_j\) be the
subgraph spanned by the internal lines whose momenta have size at least
\(\rho_j\).  Then
\[
  H_1\subset H_2\subset\cdots\subset H_N .
\]
For two indices \(i<j\), every connected component of \(H_i\) is either
contained in a connected component of \(H_j\) or disjoint from it, because
the edge set of \(H_i\) is a subset of the edge set of \(H_j\).

It remains to check that passing from connected components to 1PI
loop-containing components does not introduce a crossing overlap.  Use the
block decomposition of a connected graph into maximal one-particle-irreducible
blocks, bridges, and tree attachments.  If \(B\) is a 1PI block of a smaller
threshold graph \(H_i\), then any two lines of \(B\) lie on a common cycle in
\(H_i\), hence still lie on a common cycle in the larger graph \(H_j\).
Thus \(B\) is contained in a unique maximal 1PI block of the connected
component of \(H_j\) that contains it.  Adding further high-momentum lines can
merge old 1PI blocks, but it cannot split an old 1PI block.  Consequently two
retained 1PI loop subgraphs obtained from different thresholds are either
edge-disjoint, or one is contained in the other after the obvious inclusion
of edge sets.  Discarding components with negative superficial degree removes
vertices from this partially ordered set and therefore preserves the forest
property.

Now consider a boundary stratum of the sector at which a subset of sector
ratios tends to zero, equivalently a nested collection of high-momentum
scales is sent to infinity relative to the slower scales.  At the first
nontrivial ratio the large lines form one of the components of some
\(H_j\).  If its superficial degree is negative, it is not retained in the
divergent forest; convergence of the corresponding sector variable is one of
the power-counting conclusions in the Hepp--Zimmermann estimate below.  If its
degree is nonnegative, the corresponding 1PI loop block is one of the retained
elements of \(F_S\).  For the next larger scale separation, the lines already
represented by retained blocks are replaced by local vertices; the remaining
divergent large-line configuration is therefore a divergent component of the
quotient graph.  An induction over the ordered ratios gives a nested sequence
of retained subgraphs and quotient divergences.  This is precisely the forest
associated to the ultraviolet hierarchy in the Hepp sector.
\end{proof}

\begin{quotedtheorem}[Hepp--Zimmermann sector power counting]
\label{qthm:hepp-zimmermann-sector-power-counting}
For a massive Euclidean graph with polynomial numerator and nonexceptional
external momenta, the Zimmermann forest integrand admits a finite smooth
partition of the noncompact loop-momentum region into Hepp sectors with the
following uniform property.\footnote{This is the analytic convergence theorem
of the Bogoliubov--Parasiuk--Hepp--Zimmermann construction in momentum space;
see K.~Hepp, ``Proof of the Bogoliubov--Parasiuk theorem on renormalization,''
\emph{Commun. Math. Phys.} \textbf{2} (1966) 301--326, and
W.~Zimmermann, ``Convergence of Bogoliubov's method of renormalization in
momentum space,'' \emph{Commun. Math. Phys.} \textbf{15} (1969) 208--234.}
In each sector \(S\), after a sector-adapted linear change of loop variables,
the Jacobian is bounded above and below, the angular variables range over a
compact set, and the line denominators are uniformly comparable to monomials
in radial sector variables.  If \(\rho_j\) are the noncompact variables attached
to the divergent members and quotient divergences of the sector forest, and
\(\sigma_b\) are the remaining noncompact variables with negative
power-counting degree, then the renormalized integrand satisfies an integrable
majorant of the form
\begin{equation}
  |R_\Gamma I_\Gamma(\ell,p)|\,\dd^{D L(\Gamma)}\ell
  \le
  C_S
  \prod_j
    \rho_j^{-1}\frac{\dd\rho_j}{\rho_j}
  \prod_b
    \sigma_b^{-1}\frac{\dd\sigma_b}{\sigma_b}
  \dd\Omega_S .
  \label{eq:hepp-zimmermann-imported-sector-bound}
\end{equation}
The constant \(C_S\) is uniform in the slower variables and angular variables
inside the sector.  The statement includes the simultaneous subgraph and
quotient power counting, the adapted routing needed for disjoint branches, and
the assertion that extra Taylor operators in the global forest formula do not
worsen the sector bound.
\end{quotedtheorem}

The quoted theorem is the hard analytic input.  The preceding forest extraction
and the algebraic forest formula explain how the BPHZ subtraction terms are
organized; they do not by themselves construct the sector coordinates or prove
the uniform denominator, numerator-derivative, and Jacobian estimates in
\eqref{eq:hepp-zimmermann-imported-sector-bound}.

\begin{theorem}[BPHZ finiteness for massive Euclidean scalar graphs]
\label{thm:bphz-finiteness-massive-euclidean-scalar}
Let \(\Gamma\) be a connected one-particle-irreducible Euclidean Feynman graph
in a massive scalar theory.  Assume that all external momenta and subtraction
points are chosen away from infrared singularities, that the numerator of
every graph is polynomial in loop and external momenta, and that \(t_\gamma\)
subtracts every divergent one-particle-irreducible subgraph \(\gamma\) through
degree \(\omega(\gamma)\).  Then
\[
  \int \prod_{i=1}^{L(\Gamma)}\dd^D\ell_i\,
  R_\Gamma I_\Gamma(\ell,p)
\]
is ultraviolet finite.  The counterterm produced by each connected divergent
subgraph is a polynomial in \(p_\gamma\) of degree at most \(\omega(\gamma)\).
\end{theorem}

\begin{proof}
The proof separates the algebraic subtraction mechanism from the analytic
sector estimate.  Choose the smooth partition of unity supplied by
Theorem~\ref{qthm:hepp-zimmermann-sector-power-counting}, with one compact
region and finitely many conic Hepp sectors.
Lemma~\ref{lem:hepp-sector-forest-extraction} attaches to each sector \(S\) a
forest \(F_S\) containing the superficially divergent boundary components of
that sector.  The compact region is harmless for massive Euclidean
propagators, so the only analytic question is absolute convergence on the
noncompact sectors.

The elementary estimate is the following Taylor-remainder bound.  Let
\(\gamma\) be a connected divergent subgraph, and let
\(\rho\ge1\) scale loop momenta internal to \(\gamma\) while the external
variables \(p_\gamma\) of \(\gamma\) are kept on a slower scale.  Write
\(p_\gamma=p_\gamma^\ast+a\).  Taylor's formula with integral remainder gives
\begin{equation}
  f(p_\gamma)-t_\gamma f(p_\gamma)
  =
  \sum_{|\alpha|=\omega(\gamma)+1}
  \frac{a^\alpha}{\alpha!}
  \int_0^1
    (1-u)^{\omega(\gamma)}
    \bigl(\partial_a^\alpha f\bigr)(p_\gamma^\ast+u a)\,\dd u .
  \label{eq:bphz-taylor-integral-remainder}
\end{equation}
Here \(f\) denotes the prepared integrand of \(\gamma\), with all proper
subdivergences already replaced by their renormalized expressions or local
counterterm vertices.  A derivative with respect to an external momentum of
\(\gamma\) differentiates either a polynomial numerator or a propagator whose
momentum is affine in \(p_\gamma\).  For a propagator,
\[
  \partial_{p_\mu}
  \frac{1}{q(\ell,p)^2+m^2}
  =
  -\,\frac{2q_\nu(\ell,p)\,\partial_{p_\mu}q_\nu(\ell,p)}
  {(q(\ell,p)^2+m^2)^2},
\]
which has one less power under the large internal scaling than the original
propagator.  Derivatives of the polynomial numerator are bounded by the same
rule because differentiating a polynomial lowers its degree, and derivatives
falling on counterterm vertices only differentiate polynomials in slower
external variables.  Consequently, including the internal loop measure of
\(\gamma\),
\begin{equation}
  \dd^{D L(\gamma)}\ell_\gamma\,
  \partial_a^\alpha f(\rho\ell_\gamma,p_\gamma)
  =
  O\!\left(\rho^{\omega(\gamma)-|\alpha|}\right)
  \dd^{D L(\gamma)}\ell_\gamma .
  \label{eq:bphz-derivative-lowers-degree}
\end{equation}
For \(|\alpha|=\omega(\gamma)+1\), the isolated \(\gamma\)-scale exponent is
at most \(-1\).  In a single radial variable this would give the bound
\begin{equation}
  C\int_R^\infty \rho^{-1}\,\frac{\dd\rho}{\rho}<\infty .
  \label{eq:bphz-one-boundary-integral}
\end{equation}
If \(\gamma\) has negative superficial degree, the same estimate holds without
applying \(1-t_\gamma\), because then the exponent is already at most \(-1\).

This one-scale estimate explains why the subtraction degree
\(\omega(\gamma)\) is the right local degree.  It is not a proof of the
multiscale sector estimate: it does not choose an adapted loop basis for
overlapping or disjoint branches, does not control cancellation loci in line
momenta, and does not supply simultaneous subgraph-and-quotient exponents.

Now fix a Hepp sector \(S\).  Its forest \(F_S\) records the order in which the
Taylor remainders are interpreted: smaller Taylor operators act before larger
ones, and the counterterm produced by a smaller member becomes a local vertex
in the quotient graph.  Theorem~\ref{qthm:hepp-zimmermann-sector-power-counting}
is precisely the uniform analytic statement that this sector-local subtraction
architecture gives the integrable product majorant.  Write the divergent
forest and quotient variables in this sector as \(\rho_1,\ldots,\rho_r\).
\begin{equation}
  |R_\Gamma I_\Gamma(\ell,p)|\,\dd^{D L(\Gamma)}\ell
  \le
  C_S
  \prod_{j=1}^r
    \rho_j^{-1}\frac{\dd\rho_j}{\rho_j}
  \prod_b
    \sigma_b^{-1}\frac{\dd\sigma_b}{\sigma_b}
  \dd\Omega_S .
  \label{eq:bphz-sector-product-bound}
\end{equation}
Here the variables \(\rho_j\) include the divergent forest scales and the
divergent quotient scales, while the \(\sigma_b\) are the remaining
noncompact sector variables with negative superficial degree.  The angular
variables \(\Omega_S\) range over a compact part of the sector after the
radial ratios have been extracted.  Thus the right hand side is integrable in
every noncompact sector variable.

The algebraic reason this is the estimate of the actual forest formula, rather
than of a sector-dependent subtraction prescription, is Zimmermann's
inclusion-exclusion identity proved above.  In the sector \(S\), the boundary
strata that need Taylor remainders are labelled by the nested or disjoint
members of \(F_S\); the global forest formula is the sector-independent sum
over all compatible forests of \(\Gamma\).  The analytic theorem quoted above
supplies the remaining uniform facts that the global terms whose subgraphs do
not meet the boundary of \(S\) are extra subtractions of already integrable
remainders, that overlapping subgraphs are resolved by the sector hierarchy,
and that these extra terms obey the same product majorant.

The mass and nonexceptional-momentum assumptions remove the remaining
noncompact possibilities.  In the compact part of loop-momentum space the
integrand is smooth.  Away from compact sets, the only boundaries are the
ultraviolet scaling boundaries estimated above; massive denominators are
bounded away from zero on the real Euclidean integration contour.  Hence the
renormalized integral is absolutely ultraviolet finite.

The locality assertion is algebraic and occurs before integration.  By
definition,
\[
  C_\gamma I_\gamma
  =
  -\,t_\gamma\,\overline R_\gamma I_\gamma ,
\]
and \(t_\gamma\) is a Taylor polynomial in the external variables \(p_\gamma\)
of degree at most \(\omega(\gamma)\).  After integrating the internal
variables of \(\gamma\), or after inserting the recursively defined prepared
counterterm, this polynomial is a local vertex in \(\Gamma/\gamma\).  Thus
every connected divergent subgraph produces a counterterm polynomial of degree
at most \(\omega(\gamma)\).  In a theory whose power-counting counterterms
belong to a finite local operator space, the recursive operation remains
inside that finite space, giving perturbative renormalizability in the
finite-list sense used in the preceding chapter.
\end{proof}

\begin{remark}[Massless graphs and the BPHZL extension]
The theorem just stated is a massive, nonexceptional-momentum theorem.  Its
proof uses the fact that the only remaining noncompact regions after
subtraction are ultraviolet regions.  For a massless graph, a Taylor
subtraction at zero external momentum can change the small-loop-momentum
behavior of the integrand.  Thus the ultraviolet forest formula alone does not
control the infrared boundary in which a massless subgraph momentum goes to
zero.

The Lowenstein--Zimmermann extension, usually called BPHZL, augments the
subtraction data so that ultraviolet and infrared degrees are controlled
simultaneously.\footnote{See J.~H.~Lowenstein and W.~Zimmermann, ``On the
formulation of theories with zero mass propagators,'' \emph{Nucl. Phys. B}
\textbf{86} (1975) 77--103, doi:10.1016/0550-3213(75)90075-9; and
J.~H.~Lowenstein and W.~Zimmermann, ``The power counting theorem for Feynman
integrals with massless propagators,'' \emph{Commun. Math. Phys.} \textbf{44}
(1975) 73--86, doi:10.1007/BF01609059.  For the scalar critical model see also
J.~H.~Lowenstein and W.~Zimmermann, ``Infrared convergence of Feynman
integrals for the massless \(\phi^4\)-model,'' \emph{Commun. Math. Phys.}
\textbf{46} (1976) 105--118, doi:10.1007/BF01608491.}
In one standard formulation, each massless propagator is temporarily replaced
by an \(s\)-dependent massive propagator, so that \(s=1\) gives the desired
massless graph while another value of \(s\) supplies an auxiliary infrared
mass.  The subtraction operators are then Taylor operators in the external
momenta together with the auxiliary parameter, with degrees chosen from both
ultraviolet and infrared power counting.  The resulting forest formula is
local in the ultraviolet variables and has enough additional subtraction at
the infrared boundary to possess a distributional \(s\to1\) massless limit.

This is the momentum-subtraction counterpart of the Wilsonian statement used
in the critical chapters.  The Wilson--Fisher and Ising discussions work on a
massless critical surface, where zero-momentum Green functions and composite
insertions must be interpreted through an infrared-safe prescription:
nonexceptional external momenta, finite volume, a small mass followed by a
scaling limit, BPHZL, or an equivalent Wilsonian exact-RG construction.  The
massive BPHZ theorem above supplies the ultraviolet locality mechanism; it is
not, by itself, the full renormalization theorem for the massless critical
regime.
\end{remark}

\section{Oversubtractions and Zimmermann Normal Products}

Composite operators are renormalized by the same forest mechanism, with one
distinguished local insertion.  Let \(O(x)\) be a local monomial in the fields
and derivatives, and let
\[
  \Gamma_O
\]
be a graph with one vertex marked as the insertion of \(O\).  Divergent
subgraphs may contain the marked vertex or may be ordinary subgraphs away from
it.  The Taylor degree assigned to a subgraph containing the insertion is
chosen at least as large as the engineering degree of \(O\), and may be chosen
larger.  The resulting renormalized insertion is denoted
\[
  N_\delta[O](x),
\]
where \(\delta\) is the subtraction degree at the marked vertex.

It fixes a BPHZ subtraction scheme for insertions of \(O\).

Increasing \(\delta\) is an oversubtraction.

\begin{proposition}[Zimmermann identity for oversubtracted normal products]
\label{prop:zimmermann-normal-product-identity}
Let \(\delta'\ge\delta\).  Fix a finite basis \(\{O_A\}\) of local monomials,
with the same exact global and spacetime quantum numbers as \(O\), and with
engineering degree at most \(\delta'\), modulo total derivatives and field
relations already chosen as part of the operator basis.  For every graph with
one insertion of \(O\), the difference between the BPHZ-renormalized
insertion with subtraction degree \(\delta'\) and the one with subtraction
degree \(\delta\) is a finite sum of BPHZ-renormalized insertions of the
\(O_A\)'s:
\begin{equation}
  N_{\delta'}[O](x)
  =
  N_\delta[O](x)
  +
  \sum_A c_A\,N_{\delta'}[O_A](x).
  \label{eq:zimmermann-normal-product-identity}
\end{equation}
The coefficients \(c_A\) are finite functions of the subtraction points,
finite subtraction prescription, masses, and renormalized couplings.  They
are local scheme-change coordinates; they are not new nonlocal dynamical data.
\end{proposition}

\begin{proof}
Work graph by graph.  For a marked subgraph \(\gamma\) containing the
insertion vertex, write \(t_\gamma^{(\delta)}\) and
\(t_\gamma^{(\delta')}\) for the two Taylor operators.  Their difference is
the finite Taylor polynomial
\begin{equation}
  t_\gamma^{(\delta')}-t_\gamma^{(\delta)}
  =
  \sum_{\delta<|\alpha|\le\delta'}
  \frac{a^\alpha}{\alpha!}
  \left.\partial_a^\alpha\right|_{a=0}
  \label{eq:zimmermann-finite-taylor-difference}
\end{equation}
in the external variables of the marked subgraph, with all lower
subdivergences prepared in the same forest order as above.  After the marked
subgraph is contracted, each monomial \(a^\alpha\) is a local derivative
operator at the insertion point.  By the chosen local operator basis, the
resulting local insertion is a finite linear combination of the \(O_A\)'s.

Unmarked divergent subgraphs are subtracted in the same way in both normal
products.  Hence their forest sums cancel in the difference except insofar as
they prepare the coefficients multiplying the finite Taylor polynomial
\eqref{eq:zimmermann-finite-taylor-difference}.  The BPHZ finiteness theorem
applied to graphs with the corresponding local insertion shows that these
coefficients are finite after the remaining loop integrations.  Regrouping
the finite local insertions in the common oversubtracted scheme \(\delta'\)
gives \eqref{eq:zimmermann-normal-product-identity}.
\end{proof}

This is the BPHZ form of the operator mixing matrix used later in the
renormalized-operator chapter.

\section{Finite Parts and Subtraction Coordinates}

The Taylor operator includes a choice of subtraction point.  In addition, one
may add finite local polynomial terms compatible with the same subtraction
degree.  Changing these choices replaces \(C_\gamma\) by
\[
  C_\gamma+\Delta C_\gamma,
\]
where \(\Delta C_\gamma\) is again a local polynomial in \(p_\gamma\) of degree
at most \(\omega(\gamma)\).  Hence a change of subtraction prescription is a
local redefinition of the parameters multiplying local operators.  The
renormalized amplitudes expressed in one set of renormalized coordinates can be
re-expressed in another by a finite local change of coordinates.

The next chapter uses this freedom actively.  A family of subtraction
conditions indexed by a Euclidean momentum scale \(\mu\) gives
scale-dependent renormalized couplings.  The differential comparison of
neighboring values of \(\mu\) is the 1PI renormalization group.
Chapter~\ref{ch:vol2-wilsonian-effective-actions} compares the same
finite local freedom with Wilsonian local coordinates: the Taylor polynomial
assigned to a divergent subgraph is a local vertex, and changing its finite
part changes the corresponding coordinate on the local Wilsonian chart.
