\chapter{QCD Renormalization, Asymptotic Freedom, and Deep Inelastic Scattering}
\label{ch:volume-ii-19}
\ConventionsLorentzian


\section{QCD as a Renormalized Gauge Theory}

Quantum chromodynamics in four dimensions is Yang--Mills theory with gauge
group \(SU(N_c)\) coupled to \(N_f\) Dirac fermions in the fundamental
representation.  Write the quark fields as
\[
  q^i_{I\alpha}(x),
  \qquad
  i=1,\ldots,N_c,\quad I=1,\ldots,N_f,
\]
where \(i\) is color, \(I\) is flavor, and \(\alpha\) is a spinor index.  We
use the Hermitian-generator normalization fixed in
Chapter~\ref{ch:vol2-classical-yang-mills-matter}, specifically
\eqref{eq:ym-lie-algebra-conventions}, \eqref{eq:ym-trace-kappa}, and
\eqref{eq:ym-lagrangian-density},
\[
  \operatorname{tr}_{\square}(t^a t^b)=\delta^{ab}.
\]
The group-theoretic constants used in renormalization are defined by
\begin{equation}
  f^a{}_{cd}f^b{}_{cd}=C_A\delta^{ab},
  \qquad
  \operatorname{tr}_{R}(t_R^a t_R^b)=T_R\delta^{ab},
  \qquad
  t_R^a t_R^a=C_R\mathbf 1_R .
  \label{eq:qcd-casimir-definitions}
\end{equation}
For \(SU(N_c)\),
\[
  C_A=2N_c,\qquad
  T_F=1,\qquad
  C_F=\frac{N_c^2-1}{N_c}.
\]
These are twice the values often quoted with generators \(T^a\) normalized by
\(\operatorname{tr}_{\square}(T^aT^b)=\frac12\delta^{ab}\).  The physical
matrix-valued connection is unchanged; only the component coordinates and the
coupling coordinate are rescaled.  In the present convention the gauge action is
\(-\frac1{4g^2}\operatorname{tr}(F_{\mu\nu}F^{\mu\nu})\).

The renormalized coupling \(g(\mu)\) depends on a scale \(\mu\) and on the
renormalization scheme.  The beta function is
\[
  \beta(g):=\mu\frac{\dd g}{\dd\mu}.
\]
The first coefficient is obtained below from the background-field 1PI
effective action.  For \(N_f\) Dirac fermions in a representation \(R\), in a
mass-independent scheme the result is
\begin{equation}
  \beta(g)
  =
  -\frac{g^3}{16\pi^2}
  \left(
  \frac{11}{3}C_A-\frac{4}{3}T_R N_f
  \right)
  +O(g^5).
  \label{eq:qcd-one-loop-beta-general}
\end{equation}
Equivalently,
\begin{equation}
  \beta(g)
  =
  -\frac{g^3}{24\pi^2}
  \left(
  11N_c-2N_f
  \right)
  +O(g^5)
  \label{eq:qcd-one-loop-beta-fundamental}
\end{equation}
for fundamental quarks in this trace normalization.

\section{Ordinary Covariant-Gauge Green-Function Rules}

The covariant gauge-fixed rules used in ordinary Yang--Mills perturbation
theory are rules for time-ordered Green functions of the local variables
\(A_\mu^a,c^a,\bar c^a,q,\bar q\).  Their physical interpretation is supplied
by the BRST cohomology and by gauge-invariant operators, but the local rules
themselves are needed for renormalization and for comparing different gauges.

In this section write \(g\) for \(g_{\mathrm{YM}}\), and use the normalization
\eqref{eq:ym-lagrangian-density},
\[
  \mathcal L_{\mathrm{YM}}
  =
  -\frac1{4g^2}F^a_{\mu\nu}F^{a\mu\nu},
  \qquad
  F^a_{\mu\nu}
  =
  \partial_\mu A^a_\nu-\partial_\nu A^a_\mu
  +f^a{}_{bc}A^b_\mu A^c_\nu .
\]
This is the coupling-absorbed normalization of
Chapter~\ref{ch:vol2-classical-yang-mills-matter}:
\(D_\mu=\partial_\mu-\ii A_\mu\), and the
coupling is the coefficient in front of the curvature term.  In the Abelian
specialization it is related to the explicit-coupling QED convention by
\eqref{eq:ym-qed-curvature-rescaling}.  Thus the beta function extracted below
from the coefficient of \(F^a_{\mu\nu}F^{a\mu\nu}/g^2\) is written in
Yang--Mills field coordinates; QED wavefunction renormalization constants in
Chapter~\ref{ch:qed-renormalization-form-factors} use the explicit-coupling
field coordinate and must first be translated by the same rescaling before
being compared term by term.
The Lorenz gauge condition is
\[
  F^a(A)=\partial^\mu A_\mu^a .
\]
Under the infinitesimal gauge transformation
\[
  \delta_\zeta A^{a\mu}
  =
  \partial^\mu\zeta^a+f^a{}_{bc}A^{b\mu}\zeta^c ,
\]
the variation of \(F^a(A)\) is
\[
  \delta_\zeta F^a(A)
  =
  \partial_\mu
  \left(\partial^\mu\zeta^a+f^a{}_{bc}A^{b\mu}\zeta^c\right).
\]
Replacing the gauge parameter by the ghost \(c^a\), and pairing the gauge
condition with the antighost \(\bar c^a\), gives the Faddeev--Popov action
\begin{equation}
  S_{\mathrm{gh}}
  =
  -\int\dd^4x\,
  \bar c^a\partial_\mu
  \left(\partial^\mu c^a+f^a{}_{bc}A^{b\mu}c^c\right).
  \label{eq:qcd-ordinary-ghost-action}
\end{equation}
Equivalently, after integrating by parts and discarding the boundary term,
\[
  S_{\mathrm{gh}}
  =
  \int\dd^4x\,
  \partial_\mu\bar c^a
  \left(\partial^\mu c^a+f^a{}_{bc}A^{b\mu}c^c\right).
\]
The operator is the massless scalar kinetic operator, while the fields are
Grassmann odd.  With the momentum convention displayed in the figure below,
\begin{equation}
  \bra{\vac}
  \operatorname T\,\widehat{\bar c}^{\,a}(x)\widehat c_b(y)
  \ket{\vac}
  =
  \delta^a{}_{b}
  \int\frac{\dd^4k}{(2\pi)^4}
  \ee^{\ii k\cdot(x-y)}
  \frac{-\ii}{k^2-\ii\epsilon}.
  \label{eq:qcd-ordinary-ghost-propagator}
\end{equation}
The extra sign of a closed ghost loop is a Berezin-Gaussian fact, not an
independent rule.  Let \(K\) denote the free ghost kinetic operator and let
\(V\) denote any matrix of background insertions carrying the ghost-gluon
vertices.  In a finite-dimensional regulator, with the ordered Berezin
convention used for Grassmann Gaussians in the spinor path-integral chapter,
\[
  \frac{Z_{\rm gh}[V]}{Z_{\rm gh}[0]}
  =
  \frac{\int[\dd\bar c\,\dd c]\,
    \exp[-\bar c(K+V)c]}
       {\int[\dd\bar c\,\dd c]\,
    \exp[-\bar cKc]}
  =
  \det(1+K^{-1}V).
\]
Hence the connected \(n\)-insertion ghost cycle is the term
\[
  \operatorname{Tr}\log(1+K^{-1}V)
  =
  \sum_{n\ge1}\frac{(-1)^{n+1}}{n}
  \operatorname{Tr}\bigl((K^{-1}V)^n\bigr).
\]
A commuting complex scalar with the same quadratic operator would instead
produce \(\det(1+K^{-1}V)^{-1}\), and therefore the opposite sign for the same
closed cycle after the local vertex factors are held fixed.  Equivalently, in
the Wick-contraction expansion, closing a ghost cycle requires a cyclic
permutation of Grassmann-odd variables before the \(c\)'s and \(\bar c\)'s are
paired; this cyclic parity is the same closed-loop sign as for fermions.  The
Feynman rules below implement this determinant sign by assigning a factor
\(-1\) to each closed ghost loop.

The ghost-gluon vertex is read from the term
\[
  \int\dd^4x\,
  \partial_\mu\bar c^a\,f^a{}_{bc}A^{b\mu}c^c .
\]
If the momentum \(k\) is assigned to the antighost line as in
figure~\ref{fig:qcd-ordinary-covariant-gauge-rules}, the displayed local
vertex factor is
\begin{equation}
  V_{\bar c^a c^c A^{b\mu}}(k)
  =
  -k_\mu f^a{}_{bc}.
  \label{eq:qcd-ordinary-ghost-gluon-vertex}
\end{equation}
Changing the direction of the ghost arrow or the sign convention for incoming
momenta changes the displayed sign in the corresponding way; the invariant
object is the action \eqref{eq:qcd-ordinary-ghost-action}.  There is no local
\(\bar c\,c\,A\,A\) vertex in this Lorenz gauge, because
\(\partial^\mu D_\mu(A)\) is only linear in \(A_\mu\).

The \(B\)-field form of the gauge-fixing term is
\[
  \int\dd^4x\,
  \left(
    B^a\partial^\mu A_\mu^a+\frac{g^2\xi}{2}B^aB^a
  \right).
\]
Eliminating \(B^a\) gives
\begin{equation}
  S_{\mathrm{gf}}
  =
  -\frac1{2g^2\xi}
  \int\dd^4x\,
  (\partial^\mu A_\mu^a)(\partial^\nu A_\nu^a).
  \label{eq:qcd-ordinary-covariant-gauge-term}
\end{equation}
The quadratic gauge-field action is then inverted by
\begin{equation}
  \bra{\vac}
  \operatorname T\,\widehat A_{a\mu}(x)\widehat A_{b\nu}(y)
  \ket{\vac}
  =
  \int\frac{\dd^4k}{(2\pi)^4}
  \ee^{\ii k\cdot(x-y)}
  g^2
  \frac{-\ii\delta_{ab}}{k^2-\ii\epsilon}
  \left[
    \eta_{\mu\nu}
    +(\xi-1)\frac{k_\mu k_\nu}{k^2}
  \right].
  \label{eq:qcd-ordinary-gluon-propagator}
\end{equation}
The factor \(g^2\) is a consequence of the \(1/g^2\) multiplying the
Yang--Mills kinetic term in the connection normalization used here.

Expanding \(F^a_{\mu\nu}F^{a\mu\nu}\) gives the cubic and quartic gluon
self-interactions:
\begin{align}
  S_{AAA}
  &=
  -\frac1{2g^2}
  \int\dd^4x\,
  f^a{}_{bc}
  \bigl(\partial_\mu A^a_\nu-\partial_\nu A^a_\mu\bigr)
  A^{b\mu} A^{c\nu} ,
  \notag\\
  S_{AAAA}
  &=
  -\frac1{4g^2}
  \int\dd^4x\,
  f^a{}_{bc}f^a{}_{de}
  A^{b\mu}A^{c\nu}A^d_{\mu}A^e_{\nu}.
  \label{eq:qcd-ordinary-gluon-self-interactions}
\end{align}
These displayed action terms specify the convention before any momentum-flow
choice is made for a vertex tensor.  Figure~\ref{fig:qcd-ordinary-covariant-gauge-rules}
records the corresponding propagator and vertex conventions in a notation that
distinguishes quark, ghost, and gauge lines, and also distinguishes present
from absent local vertices; the visual distinction is used below when the
ghost-loop sign and the absent four-ghost vertex are discussed.

\begin{figure}[H]
  \centering
  \resizebox{0.98\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize},
    ghost/.style={densely dashed,line width=0.75pt},
    gluon/.style={decorate,decoration={snake,amplitude=1.2pt,segment length=5pt},
      line width=0.75pt}]
    \node[font=\small] at (0,3.35)
      {ordinary Lorenz-gauge Yang--Mills rules for Green functions};

    \begin{scope}[shift={(-5.35,1.45)}]
      \draw[ghost,-{Latex[length=2mm]}] (-1.45,0)--(1.45,0);
      \node[left] at (-1.45,0) {\(\bar c^a\)};
      \node[right] at (1.45,0) {\(c_b\)};
      \node[blue!70!black] at (0,0.58) {ghost propagator};
      \node[align=center] at (0,-0.68)
        {\(\displaystyle
        \frac{-\ii\delta^a{}_{b}}{k^2-\ii\epsilon}\)};
    \end{scope}

    \begin{scope}[shift={(0,1.45)}]
      \coordinate (v) at (0,0);
      \draw[ghost,-{Latex[length=2mm]}] (-1.55,0)--(v);
      \draw[ghost,-{Latex[length=2mm]}] (v)--(1.55,0);
      \draw[gluon] (v)--(0,1.28);
      \fill (v) circle (2pt);
      \node[left] at (-1.55,0) {\(\bar c^a\)};
      \node[right] at (1.55,0) {\(c^c\)};
      \node[above] at (0,1.30) {\(A^{b\mu}\)};
      \node[blue!70!black] at (-0.82,0.34) {\(k\)};
      \node[align=center] at (0,-0.75)
        {\(\displaystyle -k_\mu f^a{}_{bc}\)};
      \node[blue!70!black,align=center] at (1.35,0.83)
        {ghost-gluon\\vertex};
    \end{scope}

    \begin{scope}[shift={(5.25,1.45)}]
      \draw[gluon] (-1.55,0)--(1.55,0);
      \node[left] at (-1.55,0) {\(A_{a\mu}\)};
      \node[right] at (1.55,0) {\(A_{b\nu}\)};
      \node[blue!70!black] at (0,0.58) {gluon propagator};
      \node[align=center] at (0,-0.78)
        {\(\displaystyle
        \frac{-\ii g^2\delta_{ab}}{k^2-\ii\epsilon}
        \left[
        \eta_{\mu\nu}+(\xi-1)\frac{k_\mu k_\nu}{k^2}
        \right]\)};
    \end{scope}

    \begin{scope}[shift={(-4.8,-1.25)}]
      \coordinate (v) at (0,0);
      \draw[ghost] (-1.15,0)--(v)--(1.15,0);
      \draw[gluon] (v)--(-0.68,1.0);
      \draw[gluon] (v)--(0.68,1.0);
      \fill (v) circle (1.8pt);
      \draw[red!75!black,line width=1pt] (-0.58,-0.42)--(0.58,0.78);
      \draw[red!75!black,line width=1pt] (-0.58,0.78)--(0.58,-0.42);
      \node[align=center] at (0,-0.88)
        {no local \(\bar c c A A\)\\vertex in Lorenz gauge};
    \end{scope}

    \begin{scope}[shift={(0,-1.25)}]
      \coordinate (v) at (0,0);
      \foreach \ang in {90,210,330}
        \draw[gluon] (v)--++(\ang:1.08);
      \fill (v) circle (2pt);
      \node[align=center] at (0,-0.88)
        {three-gluon term\\from \(S_{AAA}\)};
    \end{scope}

    \begin{scope}[shift={(4.8,-1.25)}]
      \coordinate (v) at (0,0);
      \foreach \ang in {45,135,225,315}
        \draw[gluon] (v)--++(\ang:1.08);
      \fill (v) circle (2pt);
      \node[align=center,text width=2.5cm] at (0,-0.92)
        {four-gluon\\interaction from\\ \(S_{AAAA}\)};
    \end{scope}

    \node[align=center,orange!85!black] at (0,-2.95)
      {closed ghost loops acquire an extra factor \(-1\)};
  \end{tikzpicture}
  }
  \caption{The ordinary covariant-gauge rules following from the Lorenz
  Faddeev--Popov action.  Dashed arrows denote the ghost-number flow from
  antighost to ghost; wavy lines denote gauge fields.  The rules organize
  gauge-fixed Green functions, while physical state spaces are described by
  BRST cohomology.}
  \label{fig:qcd-ordinary-covariant-gauge-rules}
\end{figure}

\section{Background-Field Derivation of the One-Loop Beta Function}

\subsection{Background Gauge Datum and Effective Coupling Coordinate}
\label{subsec:qcd-background-gauge-effective-coupling}

The coupling in a nonabelian gauge theory is most cleanly extracted from the
coefficient of the background curvature term in the 1PI effective action.  Let
\(A_\mu=\widetilde A_\mu+a_\mu\), where \(\widetilde A_\mu\) is a background
connection and \(a_\mu\) is the quantum fluctuation.  The background covariant
derivative in a representation \(R\) is denoted
\[
  \widetilde D_\mu=\partial_\mu-\ii \widetilde A_\mu^a t_R^a ,
  \qquad
  \widetilde F_{\mu\nu}
  =
  \partial_\mu\widetilde A_\nu-\partial_\nu\widetilde A_\mu
  -\ii[\widetilde A_\mu,\widetilde A_\nu].
\]
In the adjoint representation,
\[
  (\widetilde D_\mu X)^a
  =
  \partial_\mu X^a+f^a{}_{bc}\widetilde A_\mu^bX^c .
\]
The structure constants and trace invariants in this calculation are those of
\eqref{eq:qcd-casimir-definitions}; in particular, for fundamental quarks of
\(SU(N_c)\) in the monograph's trace normalization,
\[
  C_A=2N_c,\qquad T_F=1,\qquad C_F=\frac{N_c^2-1}{N_c}.
\]
Choose the background-field gauge condition
\[
  \widetilde D^\mu a_\mu=0 .
\]
Under a background gauge transformation with parameter
\(\zeta=\zeta^a t^a\),
\[
  \delta \widetilde A_\mu=\widetilde D_\mu\zeta,\qquad
  \delta a_\mu=-\ii[a_\mu,\zeta],
\]
and the ghosts and quarks transform in the adjoint and \(R\) representations,
respectively.  Since this transformation acts linearly on the variables
integrated over in the path integral and as a connection transformation on
\(\widetilde A\), the resulting 1PI functional
\(\Gamma[\widetilde A]\) is invariant under background gauge transformations.
Its dimension-four term therefore has the form
\begin{equation}
  \Gamma[\widetilde A]
  \supset
  -\frac1{4g^2(\mu)}
  \int\dd^4x\,
  \widetilde F^a_{\mu\nu}\widetilde F^{a\mu\nu}
  \label{eq:qcd-background-effective-coupling}
\end{equation}
in Lorentzian signature, up to terms with more powers of
\(\widetilde F\), covariant derivatives, or quark fields.  The problem is
therefore reduced to computing one coefficient in a gauge-invariant
functional, rather than separately renormalizing gauge-fixed vertices.

Figure~\ref{fig:qcd-background-field-beta} displays the actual coefficient
extraction used below.  The useful object is not a separate renormalization of
each gauge-fixed vertex, but the determinant-weighted coefficient of
\(\int\widetilde F_{\mu\nu}^a\widetilde F^{a\mu\nu}\) in the background
effective action.

\begin{figure}[H]
  \centering
  \resizebox{0.98\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize},
    gluon/.style={decorate,decoration={snake,amplitude=1.2pt,segment length=5pt},
      line width=0.75pt},
    ghost/.style={densely dashed,line width=0.75pt},
    fermion/.style={line width=0.75pt}]
    \node[font=\small] at (0,3.65)
      {background-field one-loop coefficients};

    \begin{scope}[shift={(-5.85,1.18)}]
      \node[align=center] at (0,1.75)
        {external background\\fields \(\widetilde A\)};

      \draw[gluon] (-1.65,0.20)--(-0.80,0.20);
      \draw[gluon] (0.80,0.20)--(1.65,0.20);
      \node[left] at (-1.70,0.20) {\(\widetilde A\)};
      \node[right] at (1.70,0.20) {\(\widetilde A\)};
      \draw[gluon] (0,0.20) circle (0.80);
      \node at (0,0.20) {\(a_\mu\)};

      \draw[gluon] (-1.65,-1.05)--(-0.80,-1.05);
      \draw[gluon] (0.80,-1.05)--(1.65,-1.05);
      \node[left] at (-1.70,-1.05) {\(\widetilde A\)};
      \node[right] at (1.70,-1.05) {\(\widetilde A\)};
      \draw[ghost] (0,-1.05) circle (0.80);
      \draw[-{Latex[length=1.6mm]},ghost] (-0.55,-0.47) arc (135:45:0.78);
      \node at (0,-1.05) {\(c\)};

      \draw[gluon] (-1.65,-2.30)--(-0.80,-2.30);
      \draw[gluon] (0.80,-2.30)--(1.65,-2.30);
      \node[left] at (-1.70,-2.30) {\(\widetilde A\)};
      \node[right] at (1.70,-2.30) {\(\widetilde A\)};
      \draw[fermion,-{Latex[length=1.6mm]}] (0,-2.30) circle (0.80);
      \node at (0,-2.30) {\(q\)};
    \end{scope}

    \draw[thick,rounded corners=2pt,fill=gray!6] (-3.45,2.38) rectangle (6.15,-2.50);
    \draw[thick] (-3.45,1.55)--(6.15,1.55);
    \draw[thick] (-3.45,0.30)--(6.15,0.30);
    \draw[thick] (-3.45,-0.95)--(6.15,-0.95);
    \draw[thick] (-3.45,-1.75)--(6.15,-1.75);
    \draw[thick] (-0.30,2.38)--(-0.30,-2.50);
    \draw[thick] (3.05,2.38)--(3.05,-2.50);

    \node[font=\scriptsize\bfseries] at (-1.90,1.96) {determinant weight};
    \node[font=\scriptsize\bfseries] at (1.38,1.96) {coefficient in \(b\)};
    \node[font=\scriptsize\bfseries,align=center] at (4.58,1.96)
      {contribution to\\\(\beta_0=4b\)};

    \node[align=center] at (-1.90,0.93)
      {\(-\frac12\operatorname{Tr}\log D_A\)\\vector fluctuation};
    \node[align=center] at (1.38,0.93) {\(+\dfrac56 C_A\)};
    \node[align=center] at (4.58,0.93) {\(+\dfrac{10}{3}C_A\)};

    \node[align=center] at (-1.90,-0.32)
      {\(+\operatorname{Tr}\log D_{\rm gh}\)\\Faddeev--Popov ghosts};
    \node[align=center] at (1.38,-0.32) {\(+\dfrac1{12}C_A\)};
    \node[align=center] at (4.58,-0.32) {\(+\dfrac13 C_A\)};

    \node[align=center] at (-1.90,-1.35)
      {\(+N_f\operatorname{Tr}\log D_\psi\)\\Dirac quarks in \(R\)};
    \node[align=center] at (1.38,-1.35) {\(-\dfrac13T_RN_f\)};
    \node[align=center] at (4.58,-1.35) {\(-\dfrac43T_RN_f\)};

    \node[align=center,font=\scriptsize\bfseries] at (-1.90,-2.12)
      {sum};
    \node[align=center,font=\scriptsize\bfseries] at (1.38,-2.12)
      {\(\displaystyle b=\frac{11}{12}C_A-\frac13T_RN_f\)};
    \node[align=center,font=\scriptsize\bfseries] at (4.58,-2.12)
      {\(\displaystyle \beta_0=\frac{11}{3}C_A-\frac43T_RN_f\)};

    \node[align=center,text width=10.3cm] at (0,-3.18)
      {\(\displaystyle
        \Gamma^{(1)}[\widetilde A]\supset
        \frac{b}{8\pi^2}\frac{\mu^{-\epsilon}}{\epsilon}
        \int\dd^Dx\,\widetilde F^a_{\mu\nu}\widetilde F^{a\mu\nu}\),
        hence
        \(\displaystyle
        \mu\frac{\dd g}{\dd\mu}
        =-\frac{g^3}{16\pi^2}\beta_0+O(g^5)\).};
  \end{tikzpicture}
  }
  \caption{The background-field method keeps the effective action invariant
  under gauge transformations of \(\widetilde A\), so the one-loop running is
  read from the coefficient of \(\widetilde F^2\).  The table records the
  determinant weights and the coefficient \(b\) used in
  \eqref{eq:qcd-background-b-coefficient}; the usual one-loop coefficient is
  \(\beta_0=4b\).}
  \label{fig:qcd-background-field-beta}
\end{figure}

Set the background quark and ghost fields to zero and use background Feynman
gauge.  The part of the action quadratic in \(a_\mu,q,\bar q,c,\bar c\) has
the form
\begin{equation}
  S^{(2)}
  =
  \int\dd^4x\,
  \left[
    -\frac1{4g_0^2}a^\mu_a(D_A)_{a\mu,b\nu}a^\nu_b
    +\bar q D_\psi q
    +\bar c^aD_{\rm gh}^{ab}c^b
  \right],
  \label{eq:qcd-background-quadratic-action}
\end{equation}
where \(g_0\) is the bare coupling.  The operators are differential operators
whose coefficients are built from \(\widetilde D_\mu\) and
\(\widetilde F_{\mu\nu}\).  In background Feynman gauge their principal
pieces are
\[
  D_\psi=-\gamma^\mu\widetilde D_\mu,\qquad
  D_{\rm gh}=\widetilde D_\mu\widetilde D^\mu,
\]
and \(D_A\) is the adjoint vector operator obtained from the quadratic part of
\((\widetilde F+\widetilde D a-\widetilde D a+[a,a])^2\) together with
the gauge-fixing term.  The Gaussian integral gives
\begin{equation}
  \ii\Gamma^{(1)}[\widetilde A]
  =
  \log
  \left[
    \frac{(\det D_\psi)^{N_f}\det D_{\rm gh}}
    {(\det D_A)^{1/2}}
  \right].
  \label{eq:qcd-background-determinant-ratio}
\end{equation}
The numerator reflects Grassmann integration for quarks and ghosts; the
denominator is the square root from the bosonic gauge-field Gaussian.

\subsection{Constant-Background Extraction of the Logarithmic Coefficient}
\label{subsec:qcd-constant-background-log-extraction}

To extract the logarithmic coefficient, take \(\widetilde A_\mu\) to be
approximately constant inside a large box and to vanish smoothly outside.
The box size supplies an infrared scale \(L\sim\mu^{-1}\).  In the constant
region, the functional trace of a function of a differential operator reduces
to an ordinary momentum integral.  Figure~\ref{fig:qcd-constant-background-box}
separates the position-space infrared regulator from the momentum-space
large-\(|k|\) tail whose logarithm gives the ultraviolet pole:

\begin{figure}[H]
  \centering
  \resizebox{0.88\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}[shift={(-2.65,0)}]
      \draw[thick] (-2.25,-1.35) rectangle (2.25,1.35);
      \fill[blue!10] (-0.85,0) circle (1.02);
      \draw[blue!65!black,thick] (-0.85,0) circle (1.02);
      \node[blue!65!black,align=center] at (-0.85,0)
      {\(\widetilde A_\mu\simeq\text{const.}\)};
      \node[align=center] at (1.05,0.58)
        {\(\widetilde A_\mu\to0\)\\near the boundary};
      \draw[->,thick] (0.98,0.20)--(0.08,0.28);
      \draw[<->,purple!75!black,thick] (-2.25,-1.68)--(2.25,-1.68);
      \node[purple!75!black,below] at (0,-1.68)
        {\(L\sim\mu^{-1}\)};
      \node[align=center] at (0,-2.35) {position-space\\infrared regulator};
    \end{scope}

    \begin{scope}[shift={(3.05,0)}]
      \draw[->,thick] (-1.75,0)--(1.95,0) node[right] {\(k_1\)};
      \draw[->,thick] (0,-1.55)--(0,1.75) node[above] {\(k_2\)};
      \draw[densely dashed,line width=0.85pt] (0,0) circle (0.62);
      \node[below right] at (0.45,-0.44) {\(|k|=\mu\)};
      \draw[orange!85!black,line width=1.25pt,->] (0.78,0.58)
        arc[start angle=37,end angle=83,radius=0.98];
      \draw[orange!85!black,line width=1.25pt,->] (-0.92,-0.72)
        arc[start angle=218,end angle=168,radius=1.18];
      \draw[orange!85!black,line width=1.25pt,->] (1.05,-1.00)--(1.65,-1.42);
      \node[orange!85!black,align=center] at (0.95,1.16)
        {large-\(|k|\)\\tail};
      \node[orange!85!black] at (1.92,-1.48) {\(k\to\infty\)};
      \node[align=center] at (0,-2.35) {momentum-space\\UV logarithm};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The constant-background computation is a local extraction of the
  logarithmic coefficient.  The background is chosen nearly constant in a
  region of size \(L\), and the scale \(\mu\sim L^{-1}\) is used as the
  infrared cutoff in the momentum integral; the ultraviolet pole is the
  large-momentum tail of the same local trace.}
  \label{fig:qcd-constant-background-box}
\end{figure}

\begin{equation}
  \operatorname{Tr} f(D)
  =
  \int\dd^Dx
  \int\frac{\dd^Dk}{(2\pi)^D}\,
  \operatorname{tr} f(M(k)).
  \label{eq:qcd-constant-background-functional-trace}
\end{equation}
Here \(\operatorname{Tr}\) includes integration over spacetime and trace over
all indices, while \(\operatorname{tr}\) is the finite-dimensional trace over
spin, Lorentz, and color indices of the symbol \(M(k)\).

The coefficient can be extracted either by expanding the finite-dimensional
symbols in \eqref{eq:qcd-constant-background-functional-trace} or,
equivalently, by computing the logarithmic heat-kernel coefficient of the
same local operators.  The equivalence follows because both calculations
isolate the parity-even curvature-squared functional that renormalizes
\(1/g^2\).  In the symbol expansion, the only tensor average needed at fourth
order is
\begin{equation}
  \frac{k^{\nu_1}k^{\nu_2}k^{\nu_3}k^{\nu_4}}{(k^2)^4}
  \longrightarrow
  \frac1{(k^2)^2D(D+2)}
  \left(
    \delta^{\nu_1\nu_2}\delta^{\nu_3\nu_4}
    +\delta^{\nu_1\nu_3}\delta^{\nu_2\nu_4}
    +\delta^{\nu_1\nu_4}\delta^{\nu_2\nu_3}
  \right).
  \label{eq:qcd-k-four-rotational-average}
\end{equation}
The scalar integral carrying the logarithmic divergence is
\begin{equation}
  I_\epsilon(\mu)
  :=
  \int_{|k|>\mu}
  \frac{\dd^{4-\epsilon}k}{(2\pi)^{4-\epsilon}}\frac1{(k^2)^2}
  =
  \frac1{8\pi^2}\frac{\mu^{-\epsilon}}{\epsilon}
  +\hbox{finite}.
  \label{eq:qcd-basic-log-integral}
\end{equation}

\subsection{Heat-Kernel Determinant Calculation}
\label{subsec:qcd-heat-kernel-determinant-calculation}

For a self-contained evaluation, use the Euclidean logarithmic coefficient of
a Laplace-type operator
\[
  \mathcal L=-\nabla^2+E
\]
on a vector bundle \(\mathcal E\).  Let
\(\Omega_{\mu\nu}:=[\nabla_\mu,\nabla_\nu]\).  In flat space,
\begin{equation}
  \operatorname{Tr}\log\mathcal L
  \supset
  -I_\epsilon(\mu)
  \int\dd^Dx\,
  \operatorname{tr}_{\mathcal E}
  \left(
    \frac12 E^2+\frac1{12}\Omega_{\mu\nu}\Omega^{\mu\nu}
  \right).
  \label{eq:qcd-laplace-heat-coefficient}
\end{equation}
This is obtained from
\(\operatorname{Tr}\log\mathcal L=-\int_0^\infty
\frac{\dd t}{t}\operatorname{Tr}\ee^{-t\mathcal L}\) and the flat-space heat
kernel
\[
  \operatorname{Tr}\ee^{-t\mathcal L}
  \sim
  (4\pi t)^{-D/2}
  \int\dd^Dx\,\operatorname{tr}_{\mathcal E}
  \left[
    1+tE+t^2
    \left(
      \frac12 E^2+\frac1{12}\Omega_{\mu\nu}\Omega^{\mu\nu}
    \right)+\cdots
  \right].
\]
The pole in dimensional regularization is the same pole as
\eqref{eq:qcd-basic-log-integral}.

Let \(t_{\rm ad}^a\) denote the Hermitian adjoint-representation generators,
defined by
\[
  (t_{\rm ad}^a)^b{}_{c}=\ii f^b{}_{ac},
  \qquad
  \operatorname{tr}_{\rm ad}(t_{\rm ad}^a t_{\rm ad}^b)
  =
  C_A\delta^{ab}.
\]
The curvature endomorphism of the adjoint covariant derivative is the
anti-Hermitian operator
\[
  \operatorname{ad}\widetilde F_{\mu\nu}
  :=
  -\ii \widetilde F^a_{\mu\nu}t_{\rm ad}^a .
\]
Thus the adjoint trace used in the heat-kernel coefficient satisfies
\begin{equation}
  \operatorname{tr}_{\rm ad}
  \bigl(
    \operatorname{ad}\widetilde F_{\mu\nu}\,
    \operatorname{ad}\widetilde F^{\mu\nu}
  \bigr)
  =
  (-\ii)^2
  \widetilde F^a_{\mu\nu}\widetilde F^{b\mu\nu}
  \operatorname{tr}_{\rm ad}(t_{\rm ad}^a t_{\rm ad}^b)
  =
  -C_A\,
  \widetilde F^a_{\mu\nu}\widetilde F^{a\mu\nu},
  \label{eq:qcd-adjoint-trace-sign}
\end{equation}
where the sign comes from the two explicit factors of \(-\ii\).  For a matter
representation \(R\) with Hermitian generators \(t_R^a\),
\[
  \operatorname{tr}_{R}
  \bigl(
    (-\ii \widetilde F^a_{\mu\nu}t_R^a)
    (-\ii \widetilde F^{b\mu\nu}t_R^b)
  \bigr)
  =
  -T_R\,
  \widetilde F^a_{\mu\nu}\widetilde F^{a\mu\nu}.
\]

First consider one Dirac quark in \(R\).  The first-order operator is
\(\mathcal D_\psi:=\gamma^\mu\widetilde D_\mu\), and the logarithmic
part of \(\operatorname{Tr}\log\mathcal D_\psi\) is one half of that
of the Laplace-type operator \(-\mathcal D_\psi^{\,2}\).  With
\[
  -\mathcal D_\psi^{\,2}
  =
  -\widetilde D^2
  -\frac12\gamma^{\mu\nu}\Omega_{\mu\nu},
  \qquad
  \gamma^{\mu\nu}:=\frac12[\gamma^\mu,\gamma^\nu],
\]
one has
\[
  E_\psi=-\frac12\gamma^{\mu\nu}\Omega_{\mu\nu}.
\]
Using
\[
  \operatorname{tr}_{\rm spin}
  \gamma^{\mu\nu}\gamma^{\rho\sigma}
  =
  4\left(
    \delta^{\mu\sigma}\delta^{\nu\rho}
    -\delta^{\mu\rho}\delta^{\nu\sigma}
  \right),
\]
the heat-kernel density is
\[
  \operatorname{tr}_{\rm spin\otimes R}
  \left(
    \frac12E_\psi^2
    +\frac1{12}\Omega_{\mu\nu}\Omega^{\mu\nu}
  \right)
  =
  \frac23 T_R\,
  \widetilde F^a_{\mu\nu}\widetilde F^{a\mu\nu}.
\]
The extra factor \(1/2\) relating
\(\operatorname{Tr}\log\mathcal D_\psi\) to
\(\operatorname{Tr}\log(-\mathcal D_\psi^{\,2})\) gives
\begin{equation}
  \operatorname{Tr}\log D_\psi
  \supset
  -\frac13 T_R I_\epsilon(\mu)
  \int\dd^Dx\,
  \widetilde F^a_{\mu\nu}\widetilde F^{a\mu\nu}
  \label{eq:qcd-quark-determinant-f2}
\end{equation}
for one Dirac flavor.

For the Faddeev--Popov ghosts in background gauge, the relevant Laplace-type
operator is the scalar adjoint operator \(\mathcal L_{\rm gh}=-\widetilde
D^2\).  Hence \(E_{\rm gh}=0\), and
\eqref{eq:qcd-laplace-heat-coefficient} gives
\begin{equation}
  \operatorname{Tr}\log D_{\rm gh}
  \supset
  \frac1{12}C_A I_\epsilon(\mu)
  \int\dd^Dx\,
  \widetilde F^a_{\mu\nu}\widetilde F^{a\mu\nu}.
  \label{eq:qcd-ghost-determinant-f2}
\end{equation}

For the gauge fluctuation in background Feynman gauge, the Euclidean vector
operator has the form
\[
  (\mathcal L_A)_\mu{}^\nu
  =
  -\widetilde D^2\delta_\mu{}^\nu
  -2\,\operatorname{ad}(\widetilde F_\mu{}^\nu).
\]
It is a Laplace-type operator on one-forms valued in the adjoint bundle with
\[
  (E_A)_\mu{}^\nu=-2\,\operatorname{ad}(\widetilde F_\mu{}^\nu),
  \qquad
  (\Omega_{\rho\sigma})_\mu{}^\nu
  =
  \delta_\mu{}^\nu\,\operatorname{ad}(\widetilde F_{\rho\sigma}).
\]
The finite-dimensional trace is taken in \(D=4-\epsilon\) one-form
components and in the adjoint representation.  Using
\eqref{eq:qcd-adjoint-trace-sign} and the antisymmetry of
\(\widetilde F_{\mu\nu}\),
\begin{align}
  \operatorname{tr}\frac12E_A^2
  &=
  \frac12\operatorname{tr}_{\rm ad}
  \left[
    (-2\operatorname{ad}\widetilde F_\mu{}^\nu)
    (-2\operatorname{ad}\widetilde F_\nu{}^\mu)
  \right]
  =
  2C_A\,\widetilde F^a_{\mu\nu}\widetilde F^{a\mu\nu},
  \label{eq:qcd-vector-e-trace}\\
  \operatorname{tr}\frac1{12}\Omega_{\rho\sigma}\Omega^{\rho\sigma}
  &=
  \frac{D}{12}\operatorname{tr}_{\rm ad}
  \bigl(
    \operatorname{ad}\widetilde F_{\rho\sigma}
    \operatorname{ad}\widetilde F^{\rho\sigma}
  \bigr)
  =
  -\frac{D}{12}C_A\,
  \widetilde F^a_{\rho\sigma}\widetilde F^{a\rho\sigma}.
  \label{eq:qcd-vector-omega-trace-d}
\end{align}
The \(D\)-dependence in \eqref{eq:qcd-vector-omega-trace-d} has an
evanescent part.  Write
\[
  \mu^{-\epsilon}=1-\epsilon L_\mu+O(\epsilon^2),
\]
where \(L_\mu\) is the finite logarithm determined by the subtraction
convention in \eqref{eq:qcd-basic-log-integral}.  Then
\[
  (2-D)\frac{\mu^{-\epsilon}}{\epsilon}
  =
  (-2+\epsilon)\frac{\mu^{-\epsilon}}{\epsilon}
  =
  -\frac2\epsilon+1+2L_\mu+O(\epsilon),
  \qquad
  D\frac{\mu^{-\epsilon}}{\epsilon}
  =
  \frac4\epsilon-1-4L_\mu+O(\epsilon).
\]
Consequently, replacing \(D\) by \(4\) is legitimate for the pole
coefficient; the omitted \((D-4)I_\epsilon(\mu)\) part is a finite local
contribution.
Equations \eqref{eq:qcd-vector-e-trace} and
\eqref{eq:qcd-vector-omega-trace-d} therefore give the pole density
\[
  \left(
    2-\frac{4}{12}
  \right)C_A\,
  \widetilde F^a_{\mu\nu}\widetilde F^{a\mu\nu}
  =
  \frac53C_A\,
  \widetilde F^a_{\mu\nu}\widetilde F^{a\mu\nu}.
\]
The vector determinant contribution is therefore
\begin{equation}
  \operatorname{Tr}\log D_A
  \supset
  -\frac53 C_A I_\epsilon(\mu)
  \int\dd^Dx\,
  \widetilde F^a_{\mu\nu}\widetilde F^{a\mu\nu}.
  \label{eq:qcd-vector-determinant-f2}
\end{equation}

\subsection{One-Loop Coupling Flow}
\label{subsec:qcd-background-one-loop-coupling-flow}

Combining \eqref{eq:qcd-quark-determinant-f2},
\eqref{eq:qcd-ghost-determinant-f2}, and
\eqref{eq:qcd-vector-determinant-f2} with the determinant weights in
\eqref{eq:qcd-background-determinant-ratio} gives the coefficient
\begin{equation}
  b
  =
  -\frac{N_f}{3}T_R
  +
  \left[
    -\frac12\left(-\frac53\right)+\frac1{12}
  \right]C_A
  =
  \frac{11}{12}C_A-\frac{N_f}{3}T_R .
  \label{eq:qcd-background-b-coefficient}
\end{equation}
The term \(5C_A/6\) comes from the vector determinant after the bosonic
factor \(-1/2\), the term \(C_A/12\) from the Faddeev--Popov determinant, and
the term \(-N_fT_R/3\) from the \(N_f\) quark determinants.

Comparing the logarithmic part of \(\Gamma[\widetilde A]\) with
\eqref{eq:qcd-background-effective-coupling} gives
\begin{equation}
  -\frac1{4g^2(\mu)}
  =
  -\frac1{4g_0^2}
  +
  \frac{b}{8\pi^2}\frac{\mu^{-\epsilon}}{\epsilon}
  +\hbox{finite}.
  \label{eq:qcd-background-bare-renormalized-relation}
\end{equation}
In \(D=4-\epsilon\), write the dimensionless coupling as
\[
  \lambda^2(\mu)=\mu^\epsilon g^2(\mu).
\]
This is the gauge-theory analogue of the scalar convention
\(g^\epsilon=\mu^\epsilon[\lambda+\cdots]\): the square appears because the
Yang--Mills kinetic term is parameterized by \(1/g^2\).  Thus \(\lambda\) is
the dimensionless \(D\)-dimensional coupling coordinate, while \(g(\mu)\)
retains the engineering dimension \(-\epsilon/2\).
Equivalently,
\begin{equation}
  \frac1{g_0^2}
  =
  \mu^\epsilon
  \left[
    \frac1{\lambda^2(\mu)}
    +
    \frac{b}{2\pi^2}\frac1{\epsilon}
    +\hbox{finite}
    +O(\lambda^2)
  \right].
  \label{eq:qcd-background-bare-coupling-ms}
\end{equation}
The bare coupling \(g_0\) is independent of the auxiliary scale \(\mu\).
Differentiating \eqref{eq:qcd-background-bare-coupling-ms} at fixed \(g_0\)
and then taking \(\epsilon\to0\) gives
\begin{equation}
  \frac{\dd}{\dd\log\mu}\frac1{\lambda^2(\mu)}
  =
  \frac{b}{2\pi^2}+O(\lambda^2),
  \qquad
  \beta(\lambda)
  =
  -\frac{b}{4\pi^2}\lambda^3+O(\lambda^5).
  \label{eq:qcd-background-beta-lambda}
\end{equation}
Setting \(\epsilon=0\) and writing \(g\) for the four-dimensional
dimensionless coupling,
\[
  \frac{b}{4\pi^2}
  =
  \frac1{16\pi^2}
  \left(
    \frac{11}{3}C_A-\frac{4}{3}T_RN_f
  \right),
\]
which is \eqref{eq:qcd-one-loop-beta-general}.  For fundamental quarks of
\(SU(N_c)\), \(T_R=T_F=1\) and \(C_A=2N_c\), giving
\eqref{eq:qcd-one-loop-beta-fundamental}.

\section{Dimensional Transmutation}

Assume
\[
  b_0:=11N_c-2N_f>0 .
\]
At one loop,
\[
  \mu\frac{\dd g}{\dd\mu}
  =
  -\frac{b_0}{24\pi^2}g^3 .
\]
Integrating between \(\mu_0\) and \(\mu\) gives
\begin{equation}
  \frac1{g^2(\mu)}
  =
  \frac1{g^2(\mu_0)}
  +
  \frac{b_0}{12\pi^2}\log\frac{\mu}{\mu_0}.
  \label{eq:qcd-one-loop-running}
\end{equation}
Thus \(g(\mu)\) decreases at large \(\mu\).  This property is asymptotic
freedom.

The integration constant may be traded for a scale.  At one-loop accuracy,
\begin{equation}
  \Lambda
  =
  \mu\,
  \exp\left\{
  -\frac{12\pi^2}{b_0 g^2(\mu)}
  \right\}.
  \label{eq:qcd-lambda-one-loop}
\end{equation}
Higher-loop definitions multiply the right side by powers and series in
\(g(\mu)\), and the numerical value depends on the renormalization scheme.
The scheme-independent statement is that a dimensionless renormalized coupling
has been replaced, after a renormalization condition is chosen, by a
dimensionful RG-invariant scale, together with the quark masses.  This is
dimensional transmutation.

Perturbative predictions at external momentum scale \(Q\) are organized by
the small parameter \(g(Q)\) when \(Q\gg\Lambda\).  At scales comparable to
\(\Lambda\), the perturbative expansion is no longer a controlled expansion in
small coupling.  Statements about the spectrum, confinement, and the mass gap
then require nonperturbative input.

\begin{remark}[Mass-gap problem and the status of continuum Yang--Mills]
\label{rem:qcd-mass-gap-status}
For pure compact \(SU(N_c)\) Yang--Mills theory in four dimensions, the
existence of the continuum quantum field theory on \(\mathbb R^4\) and the
existence of a positive mass gap form the Yang--Mills Millennium problem; see
Remark~\ref{rem:four-dimensional-yang-mills-status}.  In this monograph the
pure-\(SU(N_c)\) continuum, vacuum, mass-gap, confinement, and spectral package
is Conjecture~\ref{conj:pure-yang-mills-central}.  The issue is not the formal
writing of the action or the perturbative construction of renormalized Green
functions.  It is the construction of a nonperturbative continuum measure, or
equivalently a Hilbert-space or algebraic quantum theory, whose correlation
functions satisfy the required structural properties and whose physical
spectrum above the vacuum has a strictly positive lower bound.

The finite-lattice theory at nonzero lattice spacing is a well-defined
statistical-mechanical system, and numerical as well as analytic evidence
strongly supports a gapped pure Yang--Mills continuum limit.  The proof of
that continuum limit and its gap in \(D=4\) remains a separate mathematical
problem.  The QCD discussion below therefore treats confinement, flux tubes,
and hadronic spectral statements as nonperturbative physical inputs or as
properties of regulated/lattice formulations, not as consequences of the
one-loop beta function.
\end{remark}

\section{Regulated Gauge-Invariant Matrix Elements}
\label{sec:qcd-regulated-gauge-invariant-matrix-elements}

The basic nonperturbative quantities of QCD are not matrix elements of isolated
colored fields.  They are limits, when such limits exist, of regulated
gauge-invariant matrix elements.  A calculation may use gauge-fixed quark and
gluon variables internally, but the object being defined must be a
color-singlet operator, a line operator, or a BRST-cohomology class with a
specified regulator and renormalization prescription.

\begin{definition}[Regulated gauge-invariant QCD matrix-element specification]
\label{def:qcd-regulated-matrix-element-specification}
A regulated gauge-invariant QCD matrix-element specification consists of:
\begin{enumerate}
  \item a regulator family \(\mathcal R_\epsilon\), such as a finite lattice,
  a gauge-covariant heat-kernel cutoff, or a gauge-fixed BV regulator together
  with its quantum master equation;
  \item a regulated operator \(\mathcal O_\epsilon\) built from local
  gauge-invariant fields, closed Wilson loops, or colored fields whose exposed
  color indices are joined by Wilson transporters along specified paths;
  \item the external states or density matrix in which the operator is
  evaluated, for example hadron states, the vacuum, a thermal state, or a
  finite-volume transfer-matrix state;
  \item ultraviolet, endpoint, cusp, and rapidity subtractions whenever the
  operator contains coincident fields, cusped Wilson lines, infinite Wilson
  lines, or lightlike separations;
  \item a topology in which the renormalized matrix elements are required to
  converge as \(\epsilon\to0\), usually a distribution topology in spacetime,
  momentum fraction, or detector angle.
\end{enumerate}
Only after these data have been supplied does a symbol such as a Wilson loop,
PDF, fragmentation function, soft function, quasi-PDF, static potential, or
energy-flow correlator denote a nonperturbative QCD quantity.  If the
continuum limit or lightcone limit is not known to exist, the same formula is a
regulated definition plus a continuum-existence hypothesis, not a theorem.
\end{definition}

The Wilson transporter is the mechanism by which separated colored insertions
become gauge-covariant before their endpoint indices are contracted.  For
example,
\[
  \bar q(x)W_{\square}(x,y)q(y),\qquad
  F^a_{\mu\nu}(x)W_{\rm ad}^{ab}(x,y)F^b_{\rho\sigma}(y),
\]
are gauge-invariant bilocals after the spinor and Lorentz indices are
contracted as specified.  Closed Wilson loops are obtained by taking the color
trace of a transporter around a closed curve.  Baryonic static-source
operators join \(N_c\) fundamental Wilson lines at a color-antisymmetric
junction.  Soft functions and parton distributions use Wilson lines whose
directions are fixed by the high-energy kinematics; their lightlike or cusped
nature is a source of additional renormalization, not an optional convention.

This rule is not limited to perturbation theory.  On a lattice, the transporter
is the ordered product of link variables and gauge invariance is exact at
nonzero lattice spacing.  In a continuum path-integral or Hilbert-space
construction, the same Wilson-line operator must be obtained as a renormalized
composite.  Perturbative Feynman graphs then compute short-distance
coefficients, anomalous dimensions, and matching kernels for these
nonperturbative matrix elements.  They do not replace the matrix elements by
colored asymptotic particles.  Energy correlators are a useful complementary
case: the detector operator is built from the gauge-invariant stress tensor,
so no Wilson line is needed in the operator definition, although perturbative
representatives of the same observable still involve colored internal
coordinates after factorization.

\section{Infrared Possibilities and Wilson Loop Diagnostics}

The one-loop beta function determines the ultraviolet direction of the flow
when \(B_0>0\).  It does not by itself determine the infrared theory.  The
first controlled alternative to strong infrared coupling is visible only after
the two-loop term is kept.  Put
\[
  a(\mu):=\frac{g(\mu)^2}{16\pi^2}.
\]
In a mass-independent gauge-invariant renormalization scheme, the two-loop
beta function has the form
\begin{equation}
  \beta_a(a)
  :=
  \mu\frac{\dd a}{\dd\mu}
  =
  -2B_0a^2-2B_1a^3+O(a^4),
  \label{eq:banks-zaks-beta-a-general}
\end{equation}
where the one-loop coefficient is the one derived above,
\[
  B_0=\frac{11}{3}C_A-\frac{4}{3}T_RN_f,
\]
and the two-loop coefficient is
\begin{equation}
  B_1
  =
  \frac{34}{3}C_A^2
  -
  \left(
    \frac{20}{3}C_AT_R+4C_RT_R
  \right)N_f .
  \label{eq:banks-zaks-two-loop-coefficient}
\end{equation}
The constants \(C_A,T_R,C_R\) are those of
\eqref{eq:qcd-casimir-definitions}; in particular no hidden
\(\operatorname{tr}(T^aT^b)=\frac12\delta^{ab}\) convention is being used.
The coefficient \(B_1\) is the universal two-loop coefficient of the
mass-independent gauge beta function.  The one-loop derivation in the
preceding section fixes \(B_0\) in the present normalization; the present
subsection uses the two-loop coefficient to derive the fixed point and its
domain of validity.

The universality of \(B_0\) and \(B_1\) is elementary at the level of coupling
coordinates.  Let \(a'\) be another dimensionless coupling coordinate with
\[
  a'=a+r_1a^2+r_2a^3+O(a^4).
\]
Then
\[
  \beta_{a'}(a')
  =
  \frac{\dd a'}{\dd a}\,\beta_a(a)
  =
  -2B_0a'^2-2B_1a'^3+O(a'^4).
\]
Thus a small two-loop zero is not an artifact of a quadratic coordinate
change.  Its precise location beyond the first nontrivial order is
scheme-dependent, because \(O(a^4)\) terms affect the next coefficient in the
small-\(B_0\) expansion.

For fundamental quarks in the trace convention of this monograph,
\[
  C_A=2N_c,\qquad T_F=1,\qquad C_F=\frac{N_c^2-1}{N_c}.
\]
Hence
\begin{equation}
  B_0
  =
  \frac{2}{3}(11N_c-2N_f),
  \qquad
  B_1
  =
  \frac{136}{3}N_c^2
  -
  \left(\frac{52}{3}N_c-\frac4{N_c}\right)N_f .
  \label{eq:banks-zaks-fundamental-B0-B1}
\end{equation}
Asymptotic freedom requires \(B_0>0\), while a perturbative infrared zero
requires \(B_1<0\).  These inequalities give the weakly coupled Banks--Zaks
window
\begin{equation}
  \frac{34N_c^3}{13N_c^2-3}
  <
  N_f
  <
  \frac{11}{2}N_c .
  \label{eq:banks-zaks-window-core}
\end{equation}
inside the larger asymptotically free range \(N_f<11N_c/2\).  More precisely,
define the distance from the upper edge by
\begin{equation}
  \epsilon_{\rm BZ}
  :=
  \frac{11N_c-2N_f}{3},
  \qquad
  N_f=\frac{11}{2}N_c-\frac32\epsilon_{\rm BZ}.
  \label{eq:banks-zaks-epsilon}
\end{equation}
Then \(B_0=2\epsilon_{\rm BZ}\), and
\[
  B_1
  =
  -2(25N_c^2-11)
  +
  \left(26N_c-\frac6{N_c}\right)\epsilon_{\rm BZ}.
\]
The two-loop zero is
\begin{equation}
  a_*^{(2)}
  =
  -\frac{B_0}{B_1}
  =
  \frac{2\epsilon_{\rm BZ}}
       {2(25N_c^2-11)
        -(26N_c-6/N_c)\epsilon_{\rm BZ}},
  \label{eq:banks-zaks-two-loop-zero}
\end{equation}
and the fixed-point coordinate of the full perturbative beta function is
\begin{equation}
  a_*
  =
  \frac{\epsilon_{\rm BZ}}{25N_c^2-11}
  +O(\epsilon_{\rm BZ}^2).
  \label{eq:banks-zaks-leading-fixed-point}
\end{equation}
The \(O(\epsilon_{\rm BZ}^2)\) coefficient is not determined by
\eqref{eq:banks-zaks-beta-a-general} alone, because it receives a contribution
from the three-loop coefficient.  The leading coefficient in
\eqref{eq:banks-zaks-leading-fixed-point}, by contrast, is scheme-independent.

The linearized exponent of the coupling perturbation is obtained without an
extra assumption:
\begin{equation}
  \omega_{\rm BZ}
  :=
  \left.\frac{\partial\beta_a}{\partial a}\right|_{a=a_*}
  =
  -\frac{2B_0^2}{B_1}+O(B_0^3)
  =
  \frac{4\epsilon_{\rm BZ}^2}{25N_c^2-11}
  +O(\epsilon_{\rm BZ}^3).
  \label{eq:banks-zaks-coupling-exponent}
\end{equation}
With the convention \(\beta_a=\dd a/\dd\log\mu\), the positive sign means that
the perturbation \(a-a_*\) decays when the flow is followed toward the
infrared.  The massless theory is essential: quark mass terms are relevant
deformations of the fixed point and must be tuned to zero.  At fixed finite
\(N_c\), \(N_f\) is an integer, so \(\epsilon_{\rm BZ}\) is not continuously
adjustable in the literal theory.  The perturbative expansion may be viewed
as an analytic expansion in the fermion-determinant power \(N_f\), or more
physically in the Veneziano regime with \(N_f/N_c\) just below \(11/2\), where
the 't Hooft coupling \(N_c a_*\) is small.

The fixed points obtained in this way are the Banks--Zaks fixed points.  The
statement is perturbative: it constructs a weak-coupling zero of the
renormalized massless beta function and gives perturbative scaling data for
gauge-invariant local operators.  It is not a derivation of confinement or of
the low-flavor infrared behavior of QCD.

\begin{controlledapproximation}[Banks--Zaks perturbative fixed point versus a conformal-window theorem]
\label{ca:qcd-banks-zaks-status}
The Banks--Zaks construction used here is the following perturbative construction,
not a theorem that every theory with flavor number in
\eqref{eq:banks-zaks-window-core} is a nonperturbative conformal field theory.
One fixes a mass-independent gauge-invariant renormalization scheme, restricts
to the massless theory, treats \(N_f\) near the upper edge of asymptotic
freedom so that \(a_*=O(\epsilon_{\rm BZ})\), and computes renormalized
correlators of gauge-invariant local or light-ray operators as formal
asymptotic series in the small fixed-point coordinate.  The first two beta
coefficients determine the leading location
\eqref{eq:banks-zaks-leading-fixed-point} and the leading irrelevant exponent
\eqref{eq:banks-zaks-coupling-exponent}; higher coefficients and finite
operator normalizations are scheme coordinates.  If a nonperturbative
regulator and scaling limit are supplied and the RG trajectory remains in the
same small-coupling neighborhood, these perturbative data are the local
coordinates of the resulting infrared theory.  Without that extra
construction they are a controlled perturbative fixed-point expansion.  The
lower edge of the nonperturbative conformal window, if it exists for the
chosen gauge group and matter content, is not determined by
\eqref{eq:banks-zaks-beta-a-general}.
\end{controlledapproximation}

Figure~\ref{fig:qcd-banks-zaks-beta-shapes} compares the two beta-function
shapes relevant here: asymptotic freedom without a weak infrared zero and the
Banks--Zaks case with a perturbative infrared fixed point.

\begin{figure}[H]
  \centering
  \resizebox{0.92\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,2.45)
      {two infrared possibilities visible in a two-loop beta function};
    \draw[-{Latex[length=2mm]}] (-0.35,0)--(6.4,0) node[below] {\(a\)};
    \draw[-{Latex[length=2mm]}] (0,-1.55)--(0,1.55) node[left] {\(\beta_a(a)\)};
    \draw[blue!70!black,very thick]
      (0,0) .. controls (1.4,-0.22) and (3.1,-0.85) .. (5.7,-1.35);
    \node[blue!70!black,fill=white,inner sep=1.2pt,align=center] at (4.55,-0.98)
      {\(\beta<0\) through\\the perturbative range};
    \draw[purple!75!black,very thick,dash pattern=on 5pt off 2.5pt]
      (0,0) .. controls (0.85,-0.35) and (1.55,-0.42) .. (2.38,0)
      .. controls (3.25,0.48) and (4.25,1.0) .. (5.45,1.24);
    \fill[red!75!black] (2.38,0) circle (2.6pt);
    \node[red!75!black,align=center] at (2.08,0.58)
      {IR fixed\\point};
    \node[purple!75!black,align=center] at (4.35,1.18)
      {Banks--Zaks\\zero};
  \end{tikzpicture}
  }
  \caption{A negative one-loop coefficient implies ultraviolet asymptotic
  freedom.  The infrared behavior depends on the subsequent terms.  The solid
  curve has no weak infrared zero, while the dashed curve has a small two-loop
  zero giving a controlled perturbative fixed point; absence of such a zero
  leaves the flow outside perturbative control in the infrared.}
  \label{fig:qcd-banks-zaks-beta-shapes}
\end{figure}

For smaller \(N_f\), perturbation theory around the ultraviolet fixed point no
longer determines the infrared limit.  The confining regime discussed below is
therefore a conditional infrared scenario, not a conclusion derived from the
one-loop beta function alone.  The condition called confinement here is a
spectral and observable hypothesis: isolated finite-energy asymptotic
particles carry no color, and physical states are created from the vacuum by
gauge-invariant operators.  Deriving this spectral statement from the
four-dimensional continuum QCD Lagrangian is an open nonperturbative problem.
Quark and gluon fields remain part of the short-distance gauge-fixed
description; the confinement statement concerns the gauge-invariant
Hilbert-space spectrum and the algebra of physical observables.

The basic nonlocal observable for this confinement diagnostic is the Wilson
line.  For an oriented path \(C\) from \(x_1\) to \(x_2\), define
\begin{equation}
  W_R(C)
  =
  P\exp\left\{
    \ii\int_C A_\mu^a(x)t_R^a\,\dd x^\mu
  \right\},
  \label{eq:qcd-wilson-line-definition}
\end{equation}
where \(R\) is a representation of the gauge group, \(t_R^a\) are its
generators, and \(P\) denotes ordering along the path.  The nonlocal operator
\[
  \bar q(x_2)\,W_R(C)\,q(x_1)
\]
is gauge invariant when the endpoints transform in conjugate representations.
It creates a color-neutral state containing a quark, an antiquark, and the
gauge flux connecting them.

Figure~\ref{fig:qcd-wilson-line-loop} fixes the Wilson-line convention used for
gauge-invariant quark bilinears and for closed Wilson-loop observables: path
ordering is along the chosen curve, and endpoint representations are conjugate.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small,align=center] at (0,2.45)
        {open Wilson line};
      \draw[thick] (-2.35,0) .. controls (-1.2,0.65) and (1.05,-0.55) .. (2.35,0.08);
      \foreach \dy/\col in {-0.25/purple!70!black,0/purple!55!black,0.25/purple!70!black} {
        \draw[\col,thick,decorate,decoration={snake,amplitude=0.85pt,segment length=7pt}]
          (-2.15,\dy) .. controls (-1.05,{0.58+\dy}) and (0.95,{-0.50+\dy}) .. (2.15,{0.08+\dy});
      }
      \fill[blue!75!black] (-2.35,0) circle (2.8pt);
      \fill[red!75!black] (2.35,0.08) circle (2.8pt);
      \node[below left,blue!75!black] at (-2.35,0) {quark};
      \node[below right,red!75!black] at (2.35,0.08) {anti-quark};
      \node[above] at (0.15,0.55) {\(C\)};
      \node[purple!75!black,align=center] at (0,-1.02)
        {gauge flux carried by \(W_R(C)\)};
    \end{scope}

    \begin{scope}[shift={(6.2,0)}]
      \node[font=\small,align=center] at (0,2.45)
        {rectangular Wilson loop};
      \draw[->,thick] (-1.7,-1.0)--(1.7,-1.0)--(1.7,1.0)--(-1.7,1.0)--(-1.7,-0.92);
      \draw[<->,blue!70!black,thick] (-1.7,-1.34)--(1.7,-1.34);
      \node[blue!70!black,below] at (0,-1.34) {\(L\)};
      \draw[<->,orange!85!black,thick] (2.05,-1.0)--(2.05,1.0);
      \node[orange!85!black,right] at (2.05,0) {\(T\)};
      \node[align=center] at (0,0)
        {Euclidean time\\evolution of\\static sources};
      \node[purple!75!black,align=center] at (0,-1.95)
        {\(\langle W_R(C)\rangle\sim\ee^{-V(L)T}\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{An open Wilson line attached to endpoint fields gives a
  gauge-invariant operator with color flux between the endpoints.  A large
  Euclidean rectangular Wilson loop extracts the static potential between
  external color sources.}
  \label{fig:qcd-wilson-line-loop}
\end{figure}

In pure Yang--Mills theory, where there are no dynamical fundamental quarks,
a fundamental flux string cannot break by pair creation of fundamental matter.
It carries a charge under the center one-form symmetry.  For a large Euclidean
rectangle \(C_{L,T}\) of spatial width \(L\) and time height \(T\),
\begin{equation}
  \langle W_R(C_{L,T})\rangle
  \sim
  \ee^{-V_R(L)T}
  \qquad (T\to\infty).
  \label{eq:qcd-wilson-loop-static-potential}
\end{equation}
An area law
\[
  \langle W_R(C_{L,T})\rangle\sim\ee^{-\gamma LT}
\]
is equivalent, in this static-source setup, to a linearly growing potential
\[
  V_R(L)\sim\gamma L .
\]
Chapter~\ref{ch:lattice-yang-mills-nonperturbative-definition} gives the
finite-cutoff version of this statement: at sufficiently small lattice
\(\beta\), the character expansion of the Wilson action converges and gives an
area bound for fundamental Wilson loops.  That theorem belongs to the
strong-coupling lattice regime.  A proof of the corresponding continuum
four-dimensional pure Yang--Mills area law would require control of the
\(\beta\to\infty\) scaling trajectory and is part of the same
nonperturbative construction problem as the mass gap, as separated in
Conjecture~\ref{conj:pure-yang-mills-central}.
In QCD with dynamical fundamental quarks, the same Wilson loop is no longer a
sharp order parameter for confinement because the flux tube can break by
quark-antiquark pair creation.  The physical spectral statement remains that
asymptotic particles are color singlets.

\section{\texorpdfstring{The 't Hooft Large-\(N_c\) Expansion}{The 't Hooft Large-Nc Expansion}}

\subsection{Planar Counting and Factorization}
\label{subsec:qcd-large-n-planar-counting-factorization}

The large-\(N_c\) expansion is a statement about a sequence of gauge theories,
not about a single theory at \(N_c=3\).  The sequence used here is
\(SU(N_c)\) Yang--Mills theory with the same trace convention as above,
\[
  \operatorname{tr}_{\square}(t^at^b)=\delta^{ab},
  \qquad
  C_A=2N_c,\qquad T_F=1,
\]
and with
\begin{equation}
  \lambda:=g^2N_c
  \label{eq:qcd-large-n-lambda-definition}
\end{equation}
held fixed as \(N_c\to\infty\).  In the common half-trace convention
\(\operatorname{tr}(T^aT^b)=\frac12\delta^{ab}\), the corresponding coupling
is \(g_{\rm ht}^2=2g^2\) and the conventional 't Hooft coupling is
\(\lambda_{\rm ht}=g_{\rm ht}^2N_c=2\lambda\).  The constant factor \(2\)
has no effect on the \(N_c\)-power counting, but it matters when comparing
formulae across conventions.

For the color-counting argument it is useful to write the gauge field as a
traceless matrix,
\[
  A_\mu=A_\mu^a t^a,\qquad (A_\mu)^i{}_i=0,
\]
and to separate the color-index contractions from the spacetime integrals,
spin tensors, gauge-fixing tensors, and momentum denominators.  The
fundamental finite-dimensional identity in the present trace normalization is
\begin{equation}
  \sum_{a=1}^{N_c^2-1}
  (t^a)^i{}_{j}(t^a)^k{}_{\ell}
  =
  \delta^i{}_{\ell}\delta^k{}_{j}
  -\frac1{N_c}\delta^i{}_{j}\delta^k{}_{\ell}.
  \label{eq:qcd-large-n-su-completeness}
\end{equation}
Here is the derivation, since this is where convention errors most often
enter.  On \(\operatorname{End}(\mathbb C^{N_c})\) use the bilinear pairing
\((X,Y)=\operatorname{tr}_{\square}(XY)\).  The matrix units have completeness
kernel \(\delta^i{}_{\ell}\delta^k{}_{j}\), which is the identity operator on
all matrices.  The one-dimensional scalar subspace is spanned by
\(\mathbf 1\), whose squared norm is \(N_c\), so the projector onto the scalar
subspace has kernel \(N_c^{-1}\delta^i{}_{j}\delta^k{}_{\ell}\).  Since
\(\{t^a\}\) is an orthonormal basis of the traceless subspace, summing
\((t^a)^i{}_j(t^a)^k{}_\ell\) gives the identity projector minus the scalar
projector, which is exactly \eqref{eq:qcd-large-n-su-completeness}.
The first term is the \(U(N_c)\) double-line propagator: one color line carries
the fundamental index and the other carries the antifundamental index.  The
second term subtracts the trace part and is explicitly lower order in
\(1/N_c\).  It is therefore included in the expansion as a subleading
insertion; it cannot increase the leading power of \(N_c\).  Ghost fields in
covariant gauges are adjoint-valued and have the same double-line color
structure, with the Grassmann sign already accounted for by the determinant
argument above.  Quark propagators carry one fundamental color line.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small] at (0,2.25) {adjoint color line};
      \draw[blue!75!black,thick] (-2.0,0.18)--(2.0,0.18);
      \draw[red!75!black,thick] (2.0,-0.18)--(-2.0,-0.18);
      \draw[-{Latex[length=2mm]},blue!75!black] (-0.8,0.42)--(0.3,0.42);
      \draw[-{Latex[length=2mm]},red!75!black] (0.8,-0.42)--(-0.3,-0.42);
      \node[blue!75!black] at (-2.25,0.18) {\(i\)};
      \node[blue!75!black] at (2.25,0.18) {\(\ell\)};
      \node[red!75!black] at (2.25,-0.18) {\(k\)};
      \node[red!75!black] at (-2.25,-0.18) {\(j\)};
      \node[align=center] at (0,-1.05)
        {\(\delta^i{}_{\ell}\delta^k{}_{j}\)\\leading term of
        \eqref{eq:qcd-large-n-su-completeness}};
    \end{scope}

    \begin{scope}[shift={(6.4,0)}]
      \node[font=\small] at (0,2.25) {planar theta graph};
      \coordinate (L) at (-1.45,0);
      \coordinate (R) at (1.45,0);
      \foreach \y/\col in {0.82/blue!65!black,0/purple!70!black,-0.82/red!70!black} {
        \draw[\col,very thick]
          (L) .. controls (-0.65,\y) and (0.65,\y) .. (R);
        \draw[\col,very thick]
          ($(L)+(0,-0.12)$) .. controls (-0.65,{\y-0.12}) and
          (0.65,{\y-0.12}) .. ($(R)+(0,-0.12)$);
      }
      \filldraw[fill=white,draw=black,thick] (L) circle (4pt);
      \filldraw[fill=white,draw=black,thick] (R) circle (4pt);
      \node[align=center] at (0,-1.55)
        {\(V=2,\ E=3,\ F=3\)\\\(N_c^{V-E+F}=N_c^2\)};
    \end{scope}

    \begin{scope}[shift={(12.8,0)}]
      \node[font=\small] at (0,2.25) {one-handle theta graph};
      \coordinate (L) at (-1.45,0);
      \coordinate (R) at (1.45,0);
      \draw[blue!65!black,very thick]
        (L) .. controls (-0.65,0.9) and (0.65,0.9) .. (R);
      \draw[blue!65!black,very thick]
        ($(L)+(0,-0.12)$) .. controls (-0.65,0.78) and
        (0.65,0.78) .. ($(R)+(0,-0.12)$);
      \draw[purple!70!black,very thick,preaction={draw=white,line width=5pt}]
        (L) .. controls (-0.85,-0.55) and (0.85,0.55) .. (R);
      \draw[purple!70!black,very thick,preaction={draw=white,line width=5pt}]
        ($(L)+(0,-0.12)$) .. controls (-0.85,-0.67) and
        (0.85,0.43) .. ($(R)+(0,-0.12)$);
      \draw[red!70!black,very thick,preaction={draw=white,line width=5pt}]
        (L) .. controls (-0.85,0.55) and (0.85,-0.55) .. (R);
      \draw[red!70!black,very thick,preaction={draw=white,line width=5pt}]
        ($(L)+(0,-0.12)$) .. controls (-0.85,0.43) and
        (0.85,-0.67) .. ($(R)+(0,-0.12)$);
      \filldraw[fill=white,draw=black,thick] (L) circle (4pt);
      \filldraw[fill=white,draw=black,thick] (R) circle (4pt);
      \node[align=center] at (0,-1.55)
        {\(V=2,\ E=3,\ F=1\)\\\(N_c^{V-E+F}=N_c^0\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The adjoint propagator is represented by two oriented color lines.
  Thickening a double-line graph gives an oriented surface.  The planar and
  one-handle theta graphs have the same number of vertices and propagators;
  the nonplanar routing has two fewer closed color faces and is therefore
  suppressed by \(N_c^{-2}\) at fixed \(\lambda=g^2N_c\).}
  \label{fig:qcd-large-n-double-line-topology}
\end{figure}

The power counting is most transparent after the gauge-fixed adjoint
quadratic form and all adjoint interaction monomials are written in the
single-trace form
\begin{equation}
  S_{\rm adj}
  =
  \frac{N_c}{\lambda}
  \int\dd^4x\,
  \operatorname{tr}_{\square}
  \left(
    \frac12\Phi K\Phi+\sum_{\alpha} c_\alpha\,\Phi_{\alpha_1}\cdots
    \Phi_{\alpha_{r_\alpha}}
  \right),
  \label{eq:qcd-large-n-single-trace-action-form}
\end{equation}
where \(\Phi\) denotes adjoint gauge, ghost, or auxiliary fields, \(K\) is the
appropriate kinetic operator including spacetime and Lorentz indices, and the
\(c_\alpha\) are \(N_c\)-independent tensors once \(\lambda\) is held fixed.
This rewriting is only a normalization of perturbative variables; it is not a
new definition of QCD.  It records the same fact already visible from
\[
  \frac1{4g^2}\operatorname{tr}_{\square}F_{\mu\nu}F^{\mu\nu}
  =
  \frac{N_c}{4\lambda}\operatorname{tr}_{\square}F_{\mu\nu}F^{\mu\nu}.
\]
With the normalization \eqref{eq:qcd-large-n-single-trace-action-form}, each
adjoint propagator contributes a factor \(\lambda/N_c\), each single-trace
adjoint vertex contributes a factor \(N_c/\lambda\), and each closed color
face contributes a factor \(N_c\) from summing an index from \(1\) to \(N_c\).
Spacetime loop integrals and Lorentz numerators multiply this color factor but
do not alter the topological exponent.

Let \(\Gamma\) be a connected adjoint ribbon graph with \(V\) single-trace
vertices, \(E\) adjoint propagators, and \(F\) closed color faces.  Thickening
the ribbons and vertices produces a compact oriented surface \(\Sigma_\Gamma\)
possibly with boundaries.  If the surface has genus \(h\) and \(b\) boundaries,
then
\begin{equation}
  \chi(\Sigma_\Gamma)
  =
  V-E+F
  =
  2-2h-b .
  \label{eq:qcd-large-n-euler-characteristic}
\end{equation}
Therefore a fixed ribbon graph contributes
\begin{equation}
  \mathcal A_\Gamma
  =
  N_c^{V-E+F}\lambda^{E-V}\,
  \mathcal I_\Gamma(\lambda,\mu,\{p\})
  =
  N_c^{2-2h-b}\lambda^{E-V}\,
  \mathcal I_\Gamma(\lambda,\mu,\{p\}),
  \label{eq:qcd-large-n-ribbon-power}
\end{equation}
up to the explicitly subleading terms generated by the
\(-N_c^{-1}\delta^i{}_j\delta^k{}_{\ell}\) part of
\eqref{eq:qcd-large-n-su-completeness}.  The factor
\(\mathcal I_\Gamma\) contains the regulated momentum integrals, tensor
contractions, signs, and scheme-dependent finite parts.  Equation
\eqref{eq:qcd-large-n-ribbon-power} is the topological content of the
't Hooft expansion: at fixed \(\lambda\), changing the genus by one suppresses
a connected contribution by \(N_c^{-2}\).

The theta graph in figure~\ref{fig:qcd-large-n-double-line-topology} gives the
elementary computation.  Both routings have \(V=2\) cubic vertices and
\(E=3\) propagators, hence both carry the same power \(\lambda^{E-V}=\lambda\).
The planar routing has \(F=3\), so
\[
  \mathcal A_{\rm planar}\sim N_c^{2}\lambda\,\mathcal I_{\rm planar}.
\]
The crossed routing thickens to a torus with \(F=1\), so
\[
  \mathcal A_{\rm one\ handle}\sim N_c^{0}\lambda\,
  \mathcal I_{\rm one\ handle}.
\]
For comparable regulated integrals at the same external data, the color
topology alone gives
\begin{equation}
  \frac{\mathcal A_{\rm one\ handle}}{\mathcal A_{\rm planar}}
  =
  O(N_c^{-2})
  \qquad (\lambda \text{ fixed}).
  \label{eq:qcd-large-n-nonplanar-suppression}
\end{equation}
This is a finite color-index statement at each fixed ribbon graph.  It does
not by itself prove convergence of the perturbation series, nor does it define
the four-dimensional continuum gauge theory nonperturbatively.

\begin{controlledapproximation}[Genus truncation as a regulated observable
statement]
\label{ca:qcd-large-n-genus-truncation-remainder-bound}
Let \(X_{N_c,\Lambda,L}(\lambda;\mathcal J)\) be a regulated color-singlet
observable: the ultraviolet regulator, spatial volume, boundary conditions,
operator/source normalization, and coupling coordinate have all been fixed.
After the overall \(N_c\)-power predicted by
\eqref{eq:qcd-large-n-ribbon-power} has been factored out, a controlled
large-\(N_c\) genus expansion through genus \(H\) is the statement that on a
declared coupling and source window
\begin{equation}
  X_{N_c,\Lambda,L}
  =
  \sum_{h=0}^{H}N_c^{-2h}X_{h,\Lambda,L}
  +
  R_{H,N_c,\Lambda,L},
  \label{eq:qcd-large-n-genus-truncated-observable}
\end{equation}
with bounds
\begin{equation}
  |X_{h,\Lambda,L}|\le M_h,
  \qquad
  |R_{H,N_c,\Lambda,L}|
  \le
  \varepsilon_H N_c^{-2H-2}.
  \label{eq:qcd-large-n-genus-truncation-bound}
\end{equation}
The ribbon-graph Euler characteristic supplies the allowed powers of
\(1/N_c\).  The coefficient bounds and the remainder bound are additional
analytic information about the regulated QCD observable.  Without them,
``planar dominance'' is a formal color-topology classification, not a
controlled approximation to a physical matrix element or correlation
function.

The regulator labels in
\eqref{eq:qcd-large-n-genus-truncated-observable} are load-bearing.  A
continuum, infinite-volume, or massless limit may be taken after the
large-\(N_c\) limit only if the bounds are uniform in the corresponding
regulator variables, or if a separate argument proves equality of the
interchanged limits:
\[
  \lim_{\Lambda,L}\lim_{N_c\to\infty}X_{N_c,\Lambda,L}
  =
  \lim_{N_c\to\infty}\lim_{\Lambda,L}X_{N_c,\Lambda,L}.
\]
Large-\(N_c\) factorization, volume reduction, baryon scaling, and QCD-string
interpretations therefore require more than the fixed-graph power count: they
require the regulated state, observable class, spectral or line-operator
data, and the uniformity needed for the limit being asserted.  Contributions
which are smaller than every power of \(1/N_c\), or phase choices which become
visible only after an infinite-volume limit, are invisible to the genus
polynomial at any fixed \(H\) but can still affect a nonperturbative
large-\(N_c\) statement.
\end{controlledapproximation}

Single-trace operator insertions are counted as boundaries.  For adjoint words
\[
  \mathcal O_r(x)
  =
  \operatorname{tr}_{\square}
  \bigl(\Phi_{r_1}(x)\cdots \Phi_{r_{\ell_r}}(x)\bigr),
\]
the connected genus-\(h\) contribution to an \(m\)-point function of
unnormalized single traces scales as
\begin{equation}
  \bigl\langle
  \mathcal O_1(x_1)\cdots \mathcal O_m(x_m)
  \bigr\rangle_{\!c,h}
  =
  O\!\left(N_c^{2-2h-m}\right),
  \label{eq:qcd-large-n-single-trace-scaling}
\end{equation}
again at fixed \(\lambda\) and fixed operator words.  If
\(\widehat{\mathcal O}_r=N_c^{-1}\mathcal O_r\), then
\[
  \bigl\langle
  \widehat{\mathcal O}_1\cdots \widehat{\mathcal O}_m
  \bigr\rangle_{\!c,h}
  =
  O\!\left(N_c^{2-2h-2m}\right).
\]
In particular,
\begin{equation}
  \bigl\langle
  \widehat{\mathcal O}_1\widehat{\mathcal O}_2
  \bigr\rangle_c
  =
  O(N_c^{-2}),
  \qquad
  \langle \widehat{\mathcal O}_1\widehat{\mathcal O}_2\rangle
  =
  \langle \widehat{\mathcal O}_1\rangle
  \langle \widehat{\mathcal O}_2\rangle+O(N_c^{-2}).
  \label{eq:qcd-large-n-factorization}
\end{equation}
This large-\(N_c\) factorization is often the most useful nonperturbative
consequence of the counting: it says that fluctuations of normalized
single-trace observables are suppressed in the planar limit, provided the
large-\(N_c\) limit of the regulated theory exists.

\subsection{Eguchi--Kawai Volume Reduction}
\label{subsec:qcd-eguchi-kawai-volume-reduction}

Large-\(N_c\) factorization suggests a further possibility: in some
center-symmetric large-\(N_c\) regimes, single-trace Wilson-loop expectation
values on a large lattice may be obtained from a reduced matrix model with
fewer sites.  Double-line counting supplies factorization of normalized
single traces; volume reduction additionally requires matching the
finite-regulator loop equations in a center-symmetric sector.

Let \(\Lambda_L=(\mathbb Z/L\mathbb Z)^d\) be a periodic hypercubic lattice,
and let \(U_\mu(x)\in SU(N_c)\) be the link from \(x\) to
\(x+\hat\mu\).  The pure-gauge Wilson measure is
\[
  \dd\nu_L(U)
  =
  Z_L^{-1}
  \exp\left\{
    bN_c\sum_{x\in\Lambda_L}\sum_{\mu<\nu}
    \operatorname{Re}\operatorname{Tr}_{\square}
    U_{\mu\nu}(x)
  \right\}
  \prod_{x,\mu}\dd U_\mu(x),
\]
where
\[
  U_{\mu\nu}(x)=
  U_\mu(x)U_\nu(x+\hat\mu)
  U_\mu(x+\hat\nu)^{-1}U_\nu(x)^{-1}.
\]
The one-site Eguchi--Kawai reduced model has \(d\) matrices
\(V_\mu\in SU(N_c)\) and measure
\begin{equation}
  \dd\nu_{\rm EK}(V)
  =
  Z_{\rm EK}^{-1}
  \exp\left\{
    bN_c\sum_{\mu<\nu}
    \operatorname{Re}\operatorname{Tr}_{\square}
    V_\mu V_\nu V_\mu^{-1}V_\nu^{-1}
  \right\}
  \prod_\mu \dd V_\mu .
  \label{eq:qcd-eguchi-kawai-reduced-measure}
\end{equation}
The parameter \(b\) is a lattice coupling coordinate.  The present discussion
uses it at fixed regulator; continuum statements require a separate scaling
limit.

A path word in the reduced model is a sequence
\[
  p=((\mu_1,\epsilon_1),\ldots,(\mu_\ell,\epsilon_\ell)),
  \qquad \epsilon_r=\pm1 ,
\]
with associated matrix
\[
  V(p)=V_{\mu_1}^{\epsilon_1}\cdots V_{\mu_\ell}^{\epsilon_\ell}
\]
and displacement vector
\begin{equation}
  n_\rho(p)=
  \sum_{r=1}^{\ell}\epsilon_r\,\delta_{\rho\mu_r}.
  \label{eq:qcd-ek-word-displacement}
\end{equation}
The normalized trace
\[
  W_{\rm red}(p)=\frac1{N_c}\operatorname{Tr}_{\square}V(p)
\]
is gauge invariant under simultaneous conjugation
\(V_\mu\mapsto \Omega V_\mu\Omega^{-1}\) even when \(n(p)\ne0\).  This is
precisely why the reduced model needs an additional center-selection rule:
in the extended lattice theory, an open path is not a Wilson-loop observable,
whereas its reduced trace would otherwise be an allowed matrix observable.

The reduced action is invariant under the independent center transformations
\begin{equation}
  V_\mu\longmapsto z_\mu V_\mu,\qquad
  z_\mu\in Z(SU(N_c))\simeq\mathbb Z_{N_c}.
  \label{eq:qcd-ek-center-transformation}
\end{equation}
Under this transformation,
\begin{equation}
  W_{\rm red}(p)
  \longmapsto
  \left(\prod_{\rho=1}^{d}z_\rho^{\,n_\rho(p)}\right)
  W_{\rm red}(p).
  \label{eq:qcd-ek-open-word-center-charge}
\end{equation}

\paragraph{Center selection for reduced open words.}
Assume the reduced finite-\(N_c\) state is invariant under
\eqref{eq:qcd-ek-center-transformation}.  If the center charge of a word is
nontrivial, namely if \(n_\rho(p)\not\equiv0\pmod{N_c}\) for at least one
\(\rho\), then
\[
  \bigl\langle W_{\rm red}(p)\bigr\rangle_{\rm EK}=0 .
\]
Consequently, if a charged reduced word has a nonzero expectation value in a
chosen large-\(N_c\) state, that state cannot implement ordinary
Eguchi--Kawai volume reduction for Wilson loops.
Choose a direction \(\rho\) for which
\(n_\rho(p)\not\equiv0\pmod{N_c}\).  There exists
\(z_\rho\in\mathbb Z_{N_c}\) such that \(z_\rho^{n_\rho(p)}\ne1\).  Invariance
of the state gives
\[
  \langle W_{\rm red}(p)\rangle_{\rm EK}
  =
  z_\rho^{n_\rho(p)}
  \langle W_{\rm red}(p)\rangle_{\rm EK},
\]
where all other \(z_\sigma\)'s are set to \(1\).  The prefactor differs from
one, hence the expectation value is zero.  If a charged expectation is
nonzero, the assumed center invariance is false.  Such an expectation is an
extra reduced-model observable with no counterpart as a closed Wilson loop in the
unreduced lattice theory, so the reduced loop equations cannot be identified
with the unreduced Wilson-loop equations without further modification.

The loop-equation form of volume reduction can now be stated cleanly.

\begin{definition}[Volume-reduction structure for Wilson loops]
\label{def:qcd-volume-reduction-datum}
A volume-reduction structure for pure large-\(N_c\) lattice Yang--Mills Wilson
loops consists of the following assumptions and regulator objects:
\begin{enumerate}
\item a sequence of finite-volume Wilson lattice measures with fixed lattice
coupling coordinate \(b\), and a sequence of reduced measures such as
\eqref{eq:qcd-eguchi-kawai-reduced-measure} or a twisted/quenched variant;
\item a class \(\mathcal C\) of closed lattice paths whose normalized
Wilson-loop expectations have large-\(N_c\) limits;
\item translation invariance of the large-volume loop expectations and
large-\(N_c\) factorization of normalized single-trace loop observables;
\item unbroken reduced center symmetry strong enough to enforce the vanishing
of all reduced open words that appear in the loop equations for
\(\mathcal C\);
\item uniqueness, in the selected center-symmetric sector, of the solution to
the limiting Schwinger--Dyson loop equations with the specified boundary and
reflection-positivity data.
\end{enumerate}
When these assumptions and regulator objects are supplied, volume reduction means
\begin{equation}
  \lim_{N_c\to\infty}\lim_{L\to\infty}
  \left\langle\frac1{N_c}\operatorname{Tr}U(C)\right\rangle_L
  =
  \lim_{N_c\to\infty}
  \left\langle\frac1{N_c}\operatorname{Tr}V(C)\right\rangle_{\rm red}
  \label{eq:qcd-volume-reduction-loop-equality}
\end{equation}
for every \(C\in\mathcal C\), where \(V(C)\) is obtained from the sequence of
directions in the lattice path \(C\).  The equality concerns single-trace
Wilson-loop coordinates at fixed regulator.  Colored local fields, baryons,
finite-\(N_c\) spectra, and continuum limits require additional definitions
and additional limiting statements.
\end{definition}

The content of this structure is not decorative.  In a Schwinger--Dyson equation
for a Wilson loop, varying a link on the loop produces terms in which the loop
splits at repeated visits to the varied link.  In the unreduced theory these
split loops may be based at different lattice sites; translation invariance
and large-\(N_c\) factorization replace their joint expectation by the product
of one-loop expectations.  In the reduced model the corresponding matrix words
are all based at the single site, and open words can appear algebraically.
The center-selection rule above is what removes the charged open words.  The
uniqueness clause in
Definition~\ref{def:qcd-volume-reduction-datum} is also load-bearing: two
different center-symmetric solutions of the same limiting loop equations would
not be distinguished by the formal loop-equation matching alone.

\begin{controlledapproximation}[Status of Eguchi--Kawai reduction]
\label{ca:qcd-eguchi-kawai-status}
The original one-site Eguchi--Kawai model is a valid reduced candidate only in
regimes where its center symmetry is not spontaneously broken in the
large-\(N_c\) limit.  At weak lattice coupling in dimensions larger than two,
the center of the original one-site model breaks; the naive model then
produces a reduced large-\(N_c\) state with charged reduced-word expectation
values rather than the desired large-volume Wilson-loop theory.
Quenched Eguchi--Kawai, twisted Eguchi--Kawai, partial reduction, and adjoint
matter with periodic boundary conditions are different attempts to stabilize
the needed center realization.  Each changes the finite-regulator construction and
therefore requires its own loop-equation and symmetry analysis.  This
monograph uses volume reduction only when the relevant center, factorization,
and uniqueness hypotheses are explicitly part of the statement.
\end{controlledapproximation}

Fundamental quarks introduce boundaries rather than handles.  A closed quark
loop carries one fundamental color sum and one flavor sum.  In the fixed-flavor
't Hooft limit \(N_f\) is held fixed, and a graph with \(b_q\) quark-loop
boundaries scales as
\begin{equation}
  N_f^{b_q}N_c^{2-2h-b_q}.
  \label{eq:qcd-large-n-quark-boundary-fixed-nf}
\end{equation}
Thus each additional closed quark loop is suppressed by one power of
\(N_c^{-1}\) relative to a gluonic planar vacuum contribution.  In the
Veneziano limit, by contrast, \(N_f/N_c\) is held fixed; the factor
\[
  N_f^{b_q}N_c^{2-2h-b_q}
  =
  \left(\frac{N_f}{N_c}\right)^{b_q}N_c^{2-2h}
\]
shows that quark-loop boundaries are no longer suppressed.  The Banks--Zaks
discussion above naturally belongs to this Veneziano scaling when
\(N_f/N_c\) is tuned close to \(11/2\).

The same counting explains the leading \(N_c\)-dependence of hadronic
interactions in a confining large-\(N_c\) theory, subject to the separate
spectral assumptions stated below.  A mesonic color-singlet operator
\(M_{IJ}=\bar q_I q_J\) has a connected \(m\)-point function of order \(N_c\)
when all insertions lie on one quark boundary.  The canonically normalized
meson source is therefore \(N_c^{-1/2}M_{IJ}\), and connected \(m\)-meson
amplitudes scale as \(N_c^{1-m/2}\).  Glueball operators normalized to have
order-one connected two-point functions have connected \(m\)-point couplings
of order \(N_c^{2-m}\).

\subsection{\texorpdfstring{Large-\(N_c\) Baryons and Spin-Flavor Coordinates}{Large-Nc Baryons and Spin-Flavor Coordinates}}
\label{subsec:qcd-large-n-baryon-spin-flavor}

Baryons require a separate order of limits and a separate normalization.  In
an \(SU(N_c)\) theory with \(N_f\) fundamental quark flavors, a local baryon
operator in a chosen spin-flavor channel has the color-antisymmetric local
form
\begin{equation}
  B_{\mathcal A}(x)
  =
  \epsilon_{i_1\cdots i_{N_c}}\,
  \mathcal A^{I_1\alpha_1,\ldots,I_{N_c}\alpha_{N_c}}\,
  q^{i_1}_{I_1\alpha_1}(x)\cdots
  q^{i_{N_c}}_{I_{N_c}\alpha_{N_c}}(x),
  \label{eq:qcd-large-n-baryon-operator}
\end{equation}
where \(i_r\) are color indices, \(I_r\) are flavor indices,
\(\alpha_r\) are spinor indices, and \(\mathcal A\) is a tensor selecting the
spin-flavor channel.  The color tensor is totally antisymmetric.  Therefore,
for baryons built from quarks in the same spatial orbital, the remaining
spin-flavor wavefunction is symmetric under exchange of the \(N_c\) quarks.
This is a kinematic consequence of Fermi statistics and the
\(\epsilon\)-contraction, not a dynamical theorem about the spectrum.

Take the large-\(N_c\) limit with \(N_f\) fixed, quark masses fixed in
\(N_c\)-independent units, and the spatial volume sent to infinity in the
usual order needed to define particle states.  If a baryon state created by
\eqref{eq:qcd-large-n-baryon-operator} exists and has a Hartree description in
which each quark moves in an \(N_c\)-independent mean field, then its leading
mass is \(O(N_c)\).  Indeed there are \(N_c\) one-body contributions, while
the two-body estimate gives
\begin{equation}
  \binom{N_c}{2}\,g^2
  =
  \frac{N_c(N_c-1)}2\,\frac{\lambda}{N_c}
  =
  \frac{\lambda}{2}(N_c-1),
  \label{eq:qcd-large-n-baryon-pair-counting}
\end{equation}
which is also \(O(N_c)\) at fixed \(\lambda\).  The Hartree statement is a
controlled large-\(N_c\) organization only after the existence of such
color-singlet baryon states has been assumed or established in a regulator.
It is not a derivation of baryons from the continuum Yang--Mills path integral.

Spin and flavor splittings are subleading in the same limit.  If the baryon is
described at low energy by a slowly rotating spin-flavor collective coordinate
with moment of inertia \(I_{\rm rot}=O(N_c)\), then the rotor contribution has
the form
\begin{equation}
  \Delta M_J
  =
  \frac{J(J+1)}{2I_{\rm rot}},
  \label{eq:qcd-large-n-baryon-rotor-splitting}
\end{equation}
so fixed-spin splittings are \(O(N_c^{-1})\).  In theories with chiral
symmetry breaking, the same scaling is compatible with a soliton description
in the mesonic effective theory: \(F_\pi^2=O(N_c)\), meson interaction
vertices have the large-\(N_c\) scaling \(g_m=O(N_c^{1-m/2})\), and a soliton
mass built from the mesonic action is \(O(N_c)\).  The spin-flavor multiplet
structure often
called contracted \(SU(2N_f)\) is the algebraic form of the large-\(N_c\)
consistency conditions on baryon matrix elements, for example on pion-baryon
scattering.  In this monograph it is used only with the displayed hypotheses:
fixed \(N_f\), a confining theory with baryon states, and a low-energy regime
where the collective-coordinate description is valid.  In the Veneziano limit
with \(N_f/N_c\) fixed, the spin-flavor representation theory and baryon
entropy grow differently and must be analyzed as a distinct limit.

For two light flavors the contracted spin-flavor algebra can be written
without appealing to a quark-model Hamiltonian.  It is the large-\(N_c\)
limit of the exact one-body spin-flavor algebra acting on the symmetric
spin-flavor part of a color-singlet baryon source.

\begin{definition}[Two-flavor spin-flavor one-body generators]
\label{def:qcd-two-flavor-spin-flavor-generators}
On the spin-flavor space
\[
  \left(\mathbb C^2_{\rm spin}\otimes\mathbb C^2_{\rm isospin}\right)^{\otimes N_c}
\]
write \(\sigma_r^i\) and \(\tau_r^a\) for Pauli matrices acting on the spin
and isospin factors of the \(r\)-th quark slot.  Define
\begin{equation}
  J^i=\sum_{r=1}^{N_c}\frac{\sigma_r^i}{2},
  \qquad
  I^a=\sum_{r=1}^{N_c}\frac{\tau_r^a}{2},
  \qquad
  G^{ia}=\sum_{r=1}^{N_c}\frac{\sigma_r^i\tau_r^a}{4}.
  \label{eq:qcd-two-flavor-spin-flavor-generators}
\end{equation}
Here \(J^i\) generates spatial spin, \(I^a\) generates isospin, and
\(G^{ia}\) is the mixed spin-isospin one-body generator.  The normalization is
fixed by the displayed Pauli matrices; it is independent of the gauge-group
trace convention used in the Yang--Mills action.
\end{definition}

\paragraph{Contraction of the two-flavor baryon spin-flavor algebra.}
The generators of Definition~\ref{def:qcd-two-flavor-spin-flavor-generators}
satisfy the exact finite-\(N_c\) commutation relations
\begin{align}
  [J^i,J^j]&=\ii\varepsilon^{ijk}J^k,
  &
  [I^a,I^b]&=\ii\varepsilon^{abc}I^c,
  &
  [J^i,I^a]&=0,
  \label{eq:qcd-spin-flavor-ji-algebra}\\
  [J^i,G^{ja}]&=\ii\varepsilon^{ijk}G^{ka},
  &
  [I^a,G^{ib}]&=\ii\varepsilon^{abc}G^{ic},
  \label{eq:qcd-spin-flavor-g-vector}\\
  [G^{ia},G^{jb}]
  &=
  \frac{\ii}{4}\delta^{ij}\varepsilon^{abc}I^c
  +
  \frac{\ii}{4}\delta^{ab}\varepsilon^{ijk}J^k .
  \label{eq:qcd-spin-flavor-g-commutator}
\end{align}
Set \(X^{ia}=G^{ia}/N_c\).  On any sequence of baryon states whose matrix
elements of \(J^i\) and \(I^a\) grow no faster than \(O(N_c)\), the
commutator of the rescaled mixed generators obeys
\begin{equation}
  [X^{ia},X^{jb}]
  =
  \frac{\ii}{4N_c^2}\delta^{ij}\varepsilon^{abc}I^c
  +
  \frac{\ii}{4N_c^2}\delta^{ab}\varepsilon^{ijk}J^k
  =
  O(N_c^{-1})
  \label{eq:qcd-contracted-su4-xx}
\end{equation}
in matrix elements, and it is \(O(N_c^{-2})\) on fixed-spin and fixed-isospin
families.  The limiting algebra has \(X^{ia}\)'s transforming as vectors
under spin and isospin but commuting among themselves.  This is the
two-flavor contracted \(SU(4)\) algebra.
Operators acting on different quark slots commute, so it is enough to compute
the one-slot algebra.  Let \(s^i=\sigma^i/2\) and \(t^a=\tau^a/2\).  The
Pauli relations give
\[
  [s^i,s^j]=\ii\varepsilon^{ijk}s^k,\qquad
  \{s^i,s^j\}=\frac12\delta^{ij},
\]
and the same formulae with \(s,i,j,k\) replaced by \(t,a,b,c\).  Since spin
and isospin matrices commute with each other,
\[
  [s^i t^a,s^j t^b]
  =
  \frac12\{s^i,s^j\}[t^a,t^b]
  +
  \frac12[s^i,s^j]\{t^a,t^b\}.
\]
Substituting the Pauli relations gives
\begin{equation}
  [s^i t^a,s^j t^b]
  =
  \frac{\ii}{4}\delta^{ij}\varepsilon^{abc}t^c
  +
  \frac{\ii}{4}\delta^{ab}\varepsilon^{ijk}s^k .
  \label{eq:qcd-one-slot-gg-commutator}
\end{equation}
Summing \eqref{eq:qcd-one-slot-gg-commutator} over the \(N_c\) quark slots
gives \eqref{eq:qcd-spin-flavor-g-commutator}.  The remaining commutators
follow in the same way from the facts that \(s^i\) transforms as a spin vector
and \(t^a\) transforms as an isospin vector.

Dividing \(G^{ia}\) by \(N_c\) gives
\[
  [X^{ia},X^{jb}]=\frac1{N_c^2}[G^{ia},G^{jb}],
\]
which is \eqref{eq:qcd-contracted-su4-xx}.  If matrix elements of \(J\) and
\(I\) are at most order \(N_c\), the right-hand side is at most order
\(N_c^{-1}\).  If the spin and isospin labels are held fixed as
\(N_c\to\infty\), it is order \(N_c^{-2}\).  The commutators with \(J\) and
\(I\) in \eqref{eq:qcd-spin-flavor-g-vector} survive the rescaling because
\([J,X]=[J,G]/N_c\) and \([I,X]=[I,G]/N_c\).  Thus the limiting algebra is a
semidirect product of spin and isospin rotations with commuting generators
\(X^{ia}\).

The expansion also gives the perturbative origin of the string language used
in the next section.  Ribbon graphs are organized by the genus of an oriented
surface, while the connected vacuum functional has the form
\[
  \log Z
  =
  \sum_{h\ge0} N_c^{2-2h}\,F_h(\lambda)
\]
as a formal expansion in \(1/N_c^2\) at fixed \(\lambda\).  A confining flux
tube, when it exists, is a different and more infrared object: it is defined
by line operators and static-source sectors.  The large-\(N_c\) topological
expansion is therefore compatible with a genus expansion for a string
description of gauge dynamics, but it is not itself a proof of the continuum
QCD string or of confinement.

\section{QCD Strings and Flux-Tube Effective Theory}

The QCD string is the long-distance effective dynamics of a chromoelectric
flux tube, defined by gauge-invariant line observables and by the spectrum in
sectors with external color sources.  In pure \(SU(N_c)\) Yang--Mills theory
there is a center one-form symmetry with charge in
\(\mathbb Z_{N_c}\).  A Wilson line in a representation \(R\) has center charge
equal to its \(N\)-ality \(k_R\in\mathbb Z_{N_c}\).  A static pair of external
sources with nonzero \(k_R\) lies in a superselection sector that cannot relax
to the vacuum by local gluonic excitations alone.  The flux tube connecting the
sources is the lowest-energy state in this sector at fixed separation.

For a Euclidean rectangle \(C_{L,T}\), the transfer-matrix interpretation of
\(\langle W_R(C_{L,T})\rangle\) gives the spectral expansion
\begin{equation}
  \langle W_R(C_{L,T})\rangle
  =
  \sum_{\alpha}
  |\langle \alpha;L,R|\mathcal O_R(L)\vac\rangle|^2
  \ee^{-E_\alpha(L)T},
  \qquad T>0,
  \label{eq:qcd-string-static-source-spectral-expansion}
\end{equation}
where \(\mathcal O_R(L)\) is a gauge-invariant operator that creates the two
static sources separated by \(L\), and \(\ket{\alpha;L,R}\) are eigenstates of
the Hamiltonian in the corresponding static-source sector.  If the lowest
energy is isolated and the overlap in
\eqref{eq:qcd-string-static-source-spectral-expansion} is nonzero, then
\[
  V_R(L)=E_0(L)
  =
  -\lim_{T\to\infty}\frac1T\log \langle W_R(C_{L,T})\rangle .
\]
The string tension in this sector is the asymptotic slope, when the limit
exists,
\begin{equation}
  \sigma_R
  :=
  \lim_{L\to\infty}\frac{E_0(L)}{L}.
  \label{eq:qcd-string-tension-definition}
\end{equation}
For representations with the same \(N\)-ality, gluon screening can change the
short-distance representation carried by the flux, so the asymptotic tension
depends on the center charge \(k_R\), not on the full microscopic
representation \(R\), after all lower-energy screening channels are included.

\subsection[Center Sectors and k-String Tensions]{Center Sectors and \(k\)-String Tensions}

The preceding definition becomes sharper when the gauge theory has no
dynamical fields of nonzero center charge.  Let the microscopic gauge group be
the simply connected group \(SU(N_c)\), and let the only dynamical fields be
adjoint fields.  The electric one-form symmetry is then \(\mathbb Z_{N_c}\).
An external source in a representation \(R\) has center charge
\[
  k_R \equiv
  \text{number of fundamental boxes of }R
  \pmod{N_c}.
\]
The static-source Hilbert space decomposes into sectors labelled by this
charge.  If dynamical matter of center charge \(q\) is present, the label is
only the quotient charge left unscreened by such matter; if the global form of
the gauge group is changed, the allowed Wilson lines and the one-form symmetry
are changed as in the line-operator classification of
Chapter~\ref{ch:global-forms-higher-form-symmetry}.  Thus a \(k\)-string is a
spectral sector of a specified gauge theory, including its global form and its
dynamical matter content, rather than an intrinsic label of a Lie algebra.

\begin{definition}[Center-sector string tension]
\label{def:qcd-center-sector-string-tension}
In pure \(SU(N_c)\) Yang--Mills theory, fix
\(k\in\{1,\ldots,N_c-1\}\).  A \(k\)-string tension is the number
\begin{equation}
  \sigma_k
  =
  \lim_{L\to\infty}
  \frac{E_k(L)-E_{\rm vac}}{L},
  \label{eq:qcd-k-string-transfer-definition}
\end{equation}
when the following objects and limits exist: a regulator with a positive
transfer matrix, a static-source or winding-flux sector of electric center
charge \(k\), the lowest energy \(E_k(L)\) in that sector for flux length
\(L\), the vacuum energy subtraction \(E_{\rm vac}\) in the same finite box,
and the thermodynamic and continuum limits in which the dimensionless ratio
\(\sigma_k/\sigma_1\) stabilizes.  Charge conjugation identifies the sectors
\(k\) and \(N_c-k\), hence
\begin{equation}
  \sigma_k=\sigma_{N_c-k}.
  \label{eq:qcd-k-string-charge-conjugation}
\end{equation}
A sector is elementary at the level of tensions only if
\begin{equation}
  \sigma_k
  <
  \min_{\substack{k_1+\cdots+k_m\equiv k \!\!\!\pmod{N_c}\\ m\ge2}}
  (\sigma_{k_1}+\cdots+\sigma_{k_m})
  \label{eq:qcd-k-string-stability-criterion}
\end{equation}
for all nontrivial decompositions into allowed unscreened center charges.  If
the inequality is saturated, the lowest state may be a multi-string threshold
rather than a single stable flux tube.
\end{definition}

The normalized ratios
\[
  R_k(N_c)=\frac{\sigma_k}{\sigma_1}
\]
are nonperturbative observables.  Two closed formulas are often useful as
comparison coordinates, but neither is a theorem of nonsupersymmetric
four-dimensional Yang--Mills theory:
\begin{align}
  R_k^{\rm Casimir}(N_c)
  &=
  \frac{k(N_c-k)}{N_c-1},
  \label{eq:qcd-k-string-casimir-ratio}\\
  R_k^{\rm sine}(N_c)
  &=
  \frac{\sin(\pi k/N_c)}{\sin(\pi/N_c)}.
  \label{eq:qcd-k-string-sine-ratio}
\end{align}
The first is the ratio of quadratic Casimirs of the rank-\(k\) antisymmetric
representation and the fundamental representation.  The second is the sine
profile that appears in particular Abelianized supersymmetric BPS-vortex
regimes, and is derived in that setting in Volume~VII.  In the present
Yang--Mills discussion, both formulas are comparison functions for lattice or
other nonperturbative determinations of \(\sigma_k\).

\paragraph{Trace-delta Casimir comparison.}
Let \(t^a\) be \(SU(N_c)\) fundamental generators normalized by
\[
  \operatorname{tr}_{\square}(t^a t^b)=\delta^{ab}.
\]
Let \(A_k=\Lambda^k\mathbb C^{N_c}\) be the rank-\(k\) antisymmetric
representation, \(1\le k\le N_c-1\).  The comparison function
\eqref{eq:qcd-k-string-casimir-ratio} comes from the following
representation-theory calculation in the same trace convention as the
Yang--Mills action.  The quadratic Casimir eigenvalues are
\begin{equation}
  C_2(A_k)
  =
  \frac{k(N_c-k)(N_c+1)}{N_c},
  \qquad
  C_2(\square)
  =
  \frac{N_c^2-1}{N_c},
  \label{eq:qcd-k-string-casimirs}
\end{equation}
and hence
\[
  \frac{C_2(A_k)}{C_2(\square)}
  =
  \frac{k(N_c-k)}{N_c-1}.
\]

The trace-delta Fierz identity in the fundamental representation is
\begin{equation}
  \sum_a (t^a)^i{}_j(t^a)^m{}_n
  =
  \delta^i{}_n\delta^m{}_j
  -
  \frac1{N_c}\delta^i{}_j\delta^m{}_n .
  \label{eq:qcd-trace-delta-fierz}
\end{equation}
Contracting \(j=n\) and summing over \(j\) gives
\[
  \sum_a (t^a t^a)^i{}_m
  =
  \left(N_c-\frac1{N_c}\right)\delta^i{}_m,
\]
so \(C_2(\square)=(N_c^2-1)/N_c\).

On \(V^{\otimes k}\), with \(V=\mathbb C^{N_c}\), the total generator is
\[
  T^a_{\rm tot}=\sum_{r=1}^k t^a_{(r)} ,
\]
where \(t^a_{(r)}\) acts on the \(r\)-th tensor factor.  Its quadratic
Casimir is
\[
  \sum_a T^a_{\rm tot}T^a_{\rm tot}
  =
  \sum_{r=1}^k\sum_a t^a_{(r)}t^a_{(r)}
  +
  2\sum_{1\le r<s\le k}\sum_a t^a_{(r)}t^a_{(s)} .
\]
By \eqref{eq:qcd-trace-delta-fierz}, the two-site operator is
\[
  \sum_a t^a_{(r)}t^a_{(s)}
  =
  P_{rs}-\frac1{N_c},
\]
where \(P_{rs}\) exchanges the two tensor factors.  On
\(\Lambda^kV\), \(P_{rs}=-1\).  Therefore
\[
  C_2(A_k)
  =
  k\,\frac{N_c^2-1}{N_c}
  +
  2\binom{k}{2}\left(-1-\frac1{N_c}\right)
  =
  \frac{k(N_c-k)(N_c+1)}{N_c}.
\]
Dividing by the fundamental value gives
\eqref{eq:qcd-k-string-casimir-ratio}.  The factor-of-two convention in the
generator trace cancels in the ratio, but it is displayed because this
chapter uses the trace-delta normalization throughout.

For \(1\le k\le N_c-1\), the sine comparison function
\eqref{eq:qcd-k-string-sine-ratio} has the charge-conjugation symmetry
\[
  R_k^{\rm sine}(N_c)=R_{N_c-k}^{\rm sine}(N_c),
\]
the strict subadditivity property
\begin{equation}
  R_{k+\ell}^{\rm sine}(N_c)
  <
  R_k^{\rm sine}(N_c)+R_\ell^{\rm sine}(N_c),
  \qquad
  k,\ell\ge1,\quad k+\ell\le N_c-1,
  \label{eq:qcd-sine-subadditivity}
\end{equation}
and the fixed-\(k\), large-\(N_c\) expansion
\begin{equation}
  R_k^{\rm sine}(N_c)
  =
  k-\frac{\pi^2}{6}\frac{k(k^2-1)}{N_c^2}
  +O(N_c^{-4}).
  \label{eq:qcd-sine-large-n}
\end{equation}
The first identity is
\(\sin(\pi(N_c-k)/N_c)=\sin(\pi k/N_c)\).  The inequality follows from
\[
  \sin a+\sin b-\sin(a+b)
  =
  4\sin\frac a2\sin\frac b2\sin\frac{a+b}{2}>0,
  \qquad
  a=\frac{\pi k}{N_c},\quad b=\frac{\pi\ell}{N_c},
\]
in the range \(0<a,b\) and \(a+b<\pi\).  The expansion is the quotient of
the Taylor series for \(\sin(\pi k/N_c)\) and \(\sin(\pi/N_c)\).  These
identities are finite comparison algebra; the dynamical question is whether a
continuum gauge theory has flux-sector transfer-matrix eigenvalues whose
ratios approach one of these comparison functions.

The Casimir comparison function obeys the same charge-conjugation symmetry,
and its fixed-\(k\) expansion begins as
\begin{equation}
  R_k^{\rm Casimir}(N_c)
  =
  k-\frac{k(k-1)}{N_c}+O(N_c^{-2}).
  \label{eq:qcd-casimir-large-n}
\end{equation}
This is the expansion of
\[
  \frac{k(N_c-k)}{N_c-1}
  =
  k\frac{1-k/N_c}{1-1/N_c}.
\]

The different \(1/N_c\) behavior in
\eqref{eq:qcd-sine-large-n} and \eqref{eq:qcd-casimir-large-n} is one reason
that \(k\)-strings are useful probes of large-\(N_c\) confinement dynamics.
For the low ranks often used in lattice comparisons, the two functions give
\[
\begin{array}{c|c|c|c}
N_c & k & R_k^{\rm Casimir} & R_k^{\rm sine}\\
\hline
4 & 2 & 4/3 & \sqrt2\\
5 & 2 & 3/2 & (1+\sqrt5)/2\\
6 & 2 & 8/5 & \sqrt3\\
6 & 3 & 9/5 & 2\\
8 & 2 & 12/7 & \sqrt{2+\sqrt2}\\
8 & 3 & 15/7 & 1+\sqrt2\\
8 & 4 & 16/7 & \sqrt{4+2\sqrt2}
\end{array}
\]
Published lattice determinations for \(SU(4)\), \(SU(5)\), \(SU(6)\), and
\(SU(8)\) determine the transfer-matrix quantity \(\sigma_k/\sigma_1\) after
continuum extrapolation, with operator-overlap, finite-volume, and
lattice-spacing errors stated.  A derivation of either comparison formula
would require an independent spectral argument inside the gauge theory.  The
improved pure-gauge computations generally place the measured ratios between
the two comparison curves; the support supplied by those measurements is for
that interpolation pattern rather than for exact Casimir scaling in
four-dimensional nonsupersymmetric Yang--Mills theory.

\begin{controlledapproximation}[\(k\)-string comparison formulas]
\label{ca:qcd-k-string-comparison-formulas}
When this monograph compares lattice \(k\)-string data to
\eqref{eq:qcd-k-string-casimir-ratio} or
\eqref{eq:qcd-k-string-sine-ratio}, the comparison means precisely this:
the lattice calculation supplies a nonperturbative estimate of
\(\sigma_k/\sigma_1\), while the two displayed functions are finite
comparison coordinates with distinct group-theoretic or supersymmetric
origins.  No conclusion about the microscopic mechanism of confinement
follows from numerical closeness alone.  A mechanism claim would require an
operator-level derivation of the relevant transfer-matrix sector and its
large-\(L\) spectrum.
\end{controlledapproximation}

\subsection{Effective String Action and the Luscher Term}

Figure~\ref{fig:qcd-string-flux-tube-worldsheet} records the static-source flux
tube as the object whose ground-state energy \(E_0(L)\) defines the string
tension \(\sigma_R\) in \eqref{eq:qcd-string-tension-definition}.

\begin{figure}[H]
  \centering
  \resizebox{0.95\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small] at (0,2.35) {static-source flux tube};
      \shade[left color=purple!18,right color=purple!18,middle color=purple!45]
        (-2.25,-0.28) .. controls (-1.0,0.03) and (1.0,0.03) .. (2.25,-0.28)
        -- (2.25,0.28) .. controls (1.0,-0.03) and (-1.0,-0.03) .. (-2.25,0.28)
        -- cycle;
      \draw[purple!70!black,thick]
        (-2.25,-0.28) .. controls (-1.0,0.03) and (1.0,0.03) .. (2.25,-0.28);
      \draw[purple!70!black,thick]
        (-2.25,0.28) .. controls (-1.0,-0.03) and (1.0,-0.03) .. (2.25,0.28);
      \fill[blue!75!black] (-2.55,0) circle (3pt);
      \fill[red!75!black] (2.55,0) circle (3pt);
      \node[below,blue!75!black] at (-2.55,-0.08) {source};
      \node[below,red!75!black] at (2.55,-0.08) {antisource};
      \draw[<->,blue!70!black,thick] (-2.55,-1.0)--(2.55,-1.0);
      \node[below] at (0,-1.0) {\(L\)};
      \node[purple!70!black,align=center] at (0,0.85)
        {energy \(E_0(L)\sim\sigma_R L\)};
    \end{scope}
    \begin{scope}[shift={(7.0,0)}]
      \node[font=\small] at (0,2.35) {worldsheet fluctuations};
      \draw[gray!55,step=0.5] (-2.2,-1.1) grid (2.2,1.1);
      \draw[-{Latex[length=2mm]}] (-2.45,-1.25)--(2.55,-1.25)
        node[right] {spatial coordinate \(s\)};
      \draw[-{Latex[length=2mm]}] (-2.45,-1.25)--(-2.45,1.35)
        node[above] {Euclidean time \(\tau\)};
      \draw[purple!75!black,very thick]
        (-2.0,-0.45) .. controls (-1.2,0.65) and (-0.2,-0.55) .. (0.7,0.25)
        .. controls (1.35,0.85) and (1.75,0.18) .. (2.0,0.62);
      \draw[purple!75!black,very thick]
        (-2.0,-0.15) .. controls (-1.25,0.95) and (-0.25,-0.25) .. (0.65,0.55)
        .. controls (1.28,1.15) and (1.75,0.48) .. (2.0,0.92);
      \node[align=center,purple!75!black] at (0,-1.85)
        {transverse fields \(X^i(\tau,s)\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The QCD string is the long-distance flux-tube sector selected by
  gauge-invariant line operators.  At large separation the lowest energy is
  extensive in the source separation, while the low-energy excitations are
  transverse fluctuations of the flux-tube worldsheet.}
  \label{fig:qcd-string-flux-tube-worldsheet}
\end{figure}

The effective string description uses QCD as its microscopic definition and
expands the flux-tube energies and matrix elements at large \(L\).  Choose
worldsheet coordinates \((\tau,s)\), with
\(0\le s\le L\), and fix static gauge so that the embedding of the worldsheet
in \(D\)-dimensional spacetime is
\[
  X^0(\tau,s)=\tau,\qquad
  X^1(\tau,s)=s,\qquad
  X^i(\tau,s)\ \ (i=2,\ldots,D-1)
\]
for the transverse fluctuations.  The leading local action compatible with
worldsheet locality, Euclidean time translation, spatial rotations transverse
to the string, and the nonlinear realization of target-space Poincare symmetry
is the area action
\begin{equation}
  S_{\mathrm{eff}}
  =
  \sigma_R
  \int\dd\tau\,\dd s\,
  \sqrt{\det\!\left(\delta_{ab}
  +\partial_aX^i\partial_bX^i\right)}
  +\mu_R\int\dd\tau
  +S_{\mathrm{higher}},
  \label{eq:qcd-string-effective-action}
\end{equation}
where \(a,b\in\{\tau,s\}\), repeated transverse indices \(i\) are summed, and
\(\mu_R\) is an endpoint energy for an open string ending on external sources.
The term \(S_{\mathrm{higher}}\) contains local operators with more derivatives
and possible boundary terms.  Expanding the square root gives
\[
  S_{\mathrm{eff}}
  =
  \sigma_RLT
  +
  \frac{\sigma_R}{2}
  \int\dd\tau\,\dd s\,
  \partial_aX^i\partial_aX^i
  +\cdots .
\]
The first quantum correction to the flux-tube energy is therefore a
worldsheet determinant.  The determinant must be renormalized as an effective
field-theory determinant: local divergent terms are absorbed into the
worldsheet area coefficient, endpoint energies, and higher local operators,
while the nonlocal finite term is a prediction.

\begin{lemma}[Free transverse determinant and the Luscher term]
\label{lem:qcd-string-luscher-term}
Assume that, at distances \(L\sqrt{\sigma_R}\gg1\), the only gapless
worldsheet fields of the open static-source flux tube are the
\(c_\perp=D-2\) transverse displacement fields \(X^i\), and that the static
sources impose Dirichlet boundary conditions on these displacements at
\(s=0,L\).  After subtracting local counterterms proportional to the
worldsheet area and to the length of the two boundaries, the Gaussian
fluctuation determinant contributes
\[
  \Delta E^{\mathrm{open}}_0(L)
  =
  -\frac{\pi c_\perp}{24L}
\]
to the long-time energy extracted from the rectangular Wilson loop.  For a
closed flux tube winding a spatial circle of length \(L\), with periodic
transverse fields and with the transverse zero mode treated as the
center-of-mass coordinate rather than as an oscillator, the corresponding
finite part is
\[
  \Delta E^{\mathrm{closed}}_0(L)
  =
  -\frac{\pi c_\perp}{6L}.
\]
\end{lemma}

\begin{proof}
The coefficient \(\sigma_R\) in the quadratic action is removed from the
oscillator frequencies by the canonical rescaling
\(Y^i=\sqrt{\sigma_R}X^i\).  For one open transverse scalar with Dirichlet
boundary conditions,
\[
  Y(\tau,s)=\sum_{n\ge1}q_n(\tau)\sin\frac{\pi ns}{L},
  \qquad
  \omega_n=\frac{\pi n}{L}.
\]
The formal zero-point energy is
\(\frac12\sum_{n\ge1}\omega_n\).  A spectral regulator gives
\[
  E^{\mathrm{open}}_{\alpha}(L)
  =
  \frac12\sum_{n\ge1}\frac{\pi n}{L}\,n^{-\alpha}
  =
  \frac{\pi}{2L}\zeta_{\mathrm R}(\alpha-1).
\]
The heat-kernel expansion of this regulated determinant contains local
divergences proportional to \(L\) and to the endpoints.  These renormalize
\(\sigma_R\) and \(\mu_R\) in
\eqref{eq:qcd-string-effective-action}.  The nonlocal finite part is the
analytic continuation at \(\alpha=0\):
\[
  E^{\mathrm{open}}_{\mathrm{fin}}(L)
  =
  \frac{\pi}{2L}\zeta_{\mathrm R}(-1)
  =
  -\frac{\pi}{24L}.
\]
Multiplication by \(c_\perp\) gives the open-string coefficient.

For one periodic transverse scalar on a circle, the nonzero modes are labelled
by \(n\in\mathbb Z\setminus\{0\}\) and have
\(\omega_n=2\pi |n|/L\).  Equivalently, for each \(n\ge1\) there are sine and
cosine oscillators of frequency \(2\pi n/L\).  The zero-point contribution is
therefore
\[
  E^{\mathrm{closed}}_{\alpha}(L)
  =
  \sum_{n\ge1}\frac{2\pi n}{L}\,n^{-\alpha}
  =
  \frac{2\pi}{L}\zeta_{\mathrm R}(\alpha-1),
\]
whose finite part at \(\alpha=0\) is \(-\pi/(6L)\).  The omitted zero mode is
the transverse center-of-mass coordinate of the closed string; it produces the
ordinary volume normalization of the state rather than an \(L^{-1}\) Casimir
energy.  This proves the closed coefficient.
\end{proof}

Consequently, under the hypotheses of
Lemma~\ref{lem:qcd-string-luscher-term}, the open-string ground-state
energy has the universal large-\(L\) form
\begin{equation}
  E_0(L)
  =
  \sigma_R L+\mu_R
  -\frac{\pi(D-2)}{24L}
  +O(L^{-3})
  \label{eq:qcd-string-open-luscher}
\end{equation}
for a long open flux tube.  For a closed flux tube winding a spatial circle,
the corresponding leading form is
\begin{equation}
  E_0^{\mathrm{closed}}(L)
  =
  \sigma_k L
  -\frac{\pi(D-2)}{6L}
  +O(L^{-3}),
  \label{eq:qcd-string-closed-luscher}
\end{equation}
again under the assumption that the only gapless worldsheet fields are the
transverse displacement modes.  The coefficients of the displayed \(1/L\)
terms are universal because they depend only on the central charge \(D-2\) of
these free transverse modes.  The higher powers in \(1/L\), the spectrum of
massive worldsheet excitations, and the validity range of the expansion are
dynamical properties of the particular gauge theory.

The Nambu--Goto square-root spectrum is a useful reference point, not the
definition of the QCD string.  If all higher-derivative and boundary
coefficients in \(S_{\mathrm{higher}}\) are set to their Nambu--Goto values,
the open and closed ground-state energies expand as
\begin{align}
  E^{\mathrm{NG,open}}_0(L)
  &=
  \sigma_R L
  \sqrt{1-\frac{\pi(D-2)}{12\sigma_RL^2}}
  \notag\\
  &=
  \sigma_R L
  -\frac{\pi(D-2)}{24L}
  -\frac{\pi^2(D-2)^2}{1152\,\sigma_RL^3}
  +O(L^{-5}),
  \label{eq:qcd-string-ng-open-expansion}\\
  E^{\mathrm{NG,closed}}_0(L)
  &=
  \sigma_R L
  \sqrt{1-\frac{\pi(D-2)}{3\sigma_RL^2}}
  \notag\\
  &=
  \sigma_R L
  -\frac{\pi(D-2)}{6L}
  -\frac{\pi^2(D-2)^2}{72\,\sigma_RL^3}
  +O(L^{-5}).
  \label{eq:qcd-string-ng-closed-expansion}
\end{align}
The equality of the \(1/L\) term with
\eqref{eq:qcd-string-open-luscher} and
\eqref{eq:qcd-string-closed-luscher} is forced by the free transverse
massless fields.  Agreement at later orders is a statement about the
effective-string matching coefficients and the nonlinearly realized
target-space Lorentz symmetry, and must be checked or imposed in the
particular effective theory being used.

\begin{controlledapproximation}[Effective string expansion of a confining flux tube]
\label{ca:qcd-effective-string-expansion}
An effective-string prediction for a flux-tube level consists of a static
source sector or closed-string winding sector, a renormalized string tension,
boundary counterterms when endpoints are present, a list of gapless and
massive worldsheet fields, a local derivative expansion for
\(S_{\mathrm{higher}}\), and a declared large-distance regime
\[
  L\sqrt{\sigma_R}\gg1,
  \qquad
  E_{\mathrm{gap}}^{\mathrm{ws}}L\gg1
\]
for every massive worldsheet excitation not retained explicitly.  The
universal Luscher term is the first finite prediction under these hypotheses.
Every later coefficient is either fixed by the worldsheet symmetry constraints
within the chosen effective theory or is a matching parameter to be determined
from the microscopic gauge theory, for example from lattice static-potential
or flux-tube spectrum data.
\end{controlledapproximation}

\subsection{Excited Flux-Tube Levels}

The same quadratic worldsheet theory gives the first approximation to the
excited flux-tube spectrum.  For an open string with fixed endpoints, the
transverse oscillator modes are labelled by
\[
  n\in\mathbb Z_{>0},\qquad i=1,\ldots,c_\perp ,
  \qquad c_\perp=D-2 .
\]
Let \(N_{n,i}\in\mathbb Z_{\ge0}\) be the occupation number of the oscillator
of frequency \(\pi n/L\).  The total oscillator level is
\begin{equation}
  N=\sum_{i=1}^{c_\perp}\sum_{n\ge1}n\,N_{n,i}.
  \label{eq:qcd-string-open-oscillator-level}
\end{equation}
The Gaussian effective string therefore predicts
\begin{equation}
  E_N^{\mathrm{open}}(L)
  =
  \sigma_RL+\mu_R
  +
  \frac{\pi}{L}
  \left(
    N-\frac{c_\perp}{24}
  \right)
  +O(L^{-3})
  \label{eq:qcd-string-open-gaussian-levels}
\end{equation}
before the \(L^{-3}\) corrections from the nonlinear action and boundary
operators are included.  The number of oscillator states at level \(N\), prior
to resolving the finite symmetry group of the particular flux-tube channel, is
the coefficient of \(q^N\) in
\begin{equation}
  Z_{\mathrm{open,osc}}(q)
  =
  \prod_{n=1}^{\infty}(1-q^n)^{-c_\perp}.
  \label{eq:qcd-string-open-oscillator-generating}
\end{equation}
For \(D=4\), where \(c_\perp=2\), the first coefficients are
\[
  1,\ 2,\ 5,\ 10,\ldots .
\]
These are not yet a complete list of lattice quantum numbers.  A flux tube
between static sources also carries parity, charge conjugation when defined,
and angular momentum around the source axis; those labels are obtained by
decomposing the oscillator states under the residual spacetime symmetries of
the static-source sector.

In four spacetime dimensions, place a static fundamental source and static
antifundamental source on the \(z\)-axis at \(z=\pm L/2\).  The continuum
symmetry group preserving this source pair is the usual cylindrical group
\(D_{\infty h}\).  Its labels are:
\[
  \Lambda=0,1,2,\ldots
  \quad\Longleftrightarrow\quad
  \Sigma,\Pi,\Delta,\Phi,\ldots ,
\]
where \(\Lambda\) is the absolute value of the angular momentum about the
source axis; a sign
\(\eta=\pm1\), denoted \(g/u\), for the combined operation which exchanges the
source and antisource and inverts about the midpoint; and, only for
\(\Lambda=0\), a reflection sign \(\epsilon=\pm\) for reflection through a
plane containing the source axis.  At finite lattice spacing with the axis
chosen along a lattice direction, \(D_{\infty h}\) is reduced to the
corresponding lattice subgroup.  The continuum labels are therefore continuum
and large-distance assignments obtained after subduction and extrapolation,
not exact labels of a single finite lattice calculation.

\paragraph{Open flux-tube oscillator labels.}
Assume the Gaussian open-string sector in \(D=4\), with circular transverse
oscillators \(a_{n,+}^{\dagger}\) and \(a_{n,-}^{\dagger}\) of frequency
\(\pi n/L\), where \(n\geq1\) and the signs denote angular momentum
\(\pm1\) about the source axis.  For a Fock monomial
\[
  \prod_{n\geq1}
  \bigl(a_{n,+}^{\dagger}\bigr)^{N_{n,+}}
  \bigl(a_{n,-}^{\dagger}\bigr)^{N_{n,-}}\vac_{\rm str},
\]
the oscillator level, axial angular-momentum label, and \(g/u\) sign are
\begin{align}
  N&=\sum_{n\geq1}n(N_{n,+}+N_{n,-}),\\
  \Lambda&=\left|\sum_{n\geq1}(N_{n,+}-N_{n,-})\right|,\\
  \eta&=(-1)^{\sum_{n\geq1}n(N_{n,+}+N_{n,-})}.
  \label{eq:qcd-string-open-channel-bookkeeping}
\end{align}
For \(\Lambda=0\), reflection through a plane containing the source axis
exchanges \(+\) and \(-\) oscillators, so reflection eigenstates are the even
and odd combinations of each reflection orbit; if the monomial is pointwise
fixed by this exchange, only the even \(\Sigma^+\) state is obtained.

This is bookkeeping in the Gaussian reference Hilbert space.
The level formula is the definition of the open-string oscillator level in
\eqref{eq:qcd-string-open-oscillator-level}.  A circular transverse
oscillator is built from
\[
  X^{+}=X^2+iX^3,\qquad X^{-}=X^2-iX^3 .
\]
Under a rotation by angle \(\alpha\) about the source axis,
\(X^{\pm}\mapsto \ee^{\pm i\alpha}X^{\pm}\).  A monomial with occupations
\(N_{n,\pm}\) therefore has axial angular momentum
\(\sum_n(N_{n,+}-N_{n,-})\), and the irreducible cylindrical label records
its absolute value \(\Lambda\).

The operation defining \(g/u\) exchanges the endpoints and sends
\((z,X^2,X^3)\mapsto(-z,-X^2,-X^3)\) in the static-source coordinate system.
The mode function transforms as
\[
  \sin\frac{\pi n(L-s)}{L}
  =
  (-1)^{n+1}\sin\frac{\pi ns}{L},
\]
while the transverse vector itself contributes the additional sign \(-1\).
Thus a single \(n\)-th transverse oscillator has eigenvalue \((-1)^n\) under
the combined operation, and multiplication over all occupied oscillators gives
the sign \(\eta\) in
\eqref{eq:qcd-string-open-channel-bookkeeping}.  Reflection through a plane
containing the axis complex-conjugates \(X^+\) and \(X^-\), hence exchanges
the two circular polarizations.  For \(\Lambda>0\) this reflection relates the
two helicity components of the same two-dimensional real representation.  For
\(\Lambda=0\) it can be diagonalized, giving \(\epsilon=\pm\) when the
reflection orbit has two monomials and only the even state when the monomial
is fixed.

For the first oscillator levels in \(D=4\) this gives
\begin{center}
\begin{tabular}{c|c|c}
level \(N\) & \(D_{\infty h}\) content of the Gaussian reference sector
& total oscillator degeneracy\\
\hline
0 & \(\Sigma_g^+\) & \(1\)\\
1 & \(\Pi_u\) & \(2\)\\
2 & \(\Sigma_g^+\oplus\Pi_g\oplus\Delta_g\) & \(5\)\\
3 &
\(\Sigma_u^+\oplus\Sigma_u^-\oplus2\Pi_u\oplus\Delta_u\oplus\Phi_u\)
& \(10\)
\end{tabular}
\end{center}
At the next level the decomposition contains
\[
  3\Sigma_g^+\oplus\Sigma_g^-
  \oplus3\Pi_g\oplus3\Delta_g\oplus\Phi_g\oplus(\Lambda=4)_g,
\]
with total oscillator degeneracy \(20\).  Thus \(\Sigma_g^-\) is an allowed
cylindrical channel, but it is not present in the first three free
transverse-oscillator levels.  This distinction matters when one compares the
long-string reference spectrum with hybrid static-potential data.

The table is a long-string reference decomposition.  A measured hybrid static
potential in the same \(\Lambda_\eta^\epsilon\) channel may contain massive
surface excitations, endpoint-local excitations, or short-distance gluelump
physics that are not present in the free transverse oscillator Hilbert space.
Conversely, the free oscillator labels become useful at large \(L\) only
after a transfer-matrix calculation has isolated the corresponding static
sector and after the effective-string expansion is within its declared
regime.

In the lattice program initiated by Juge, Kuti, and Morningstar, the objects
called \(\Pi_u\), \(\Sigma_g^-\), \(\Sigma_u^\pm\), \(\Pi_g\), and so on are
channel-resolved static energies extracted from correlation matrices of
gauge-invariant source operators subduced to the finite lattice subgroup of
\(D_{\infty h}\).  The label is therefore assigned before an effective-string
fit is made.  At small separation the same channel can be governed by a
gluelump-like excitation of the gauge field bound to a compact static
color-octet source pair; at large separation, if no lower massive worldsheet
or endpoint excitation intervenes, the channel may approach one of the
long-string oscillator towers above.  The observed \(L\)-dependence of a
hybrid static potential is the interpolation between these two spectral
descriptions, with the Nambu--Goto reference formula supplying only the
asymptotic massless-transverse-oscillator comparison.
The channels \(\Pi_u\) and \(\Sigma_g^-\) are especially useful diagnostics:
\(\Pi_u\) is the first transverse oscillator tower in the Gaussian
long-string limit, whereas a low-lying \(\Sigma_g^-\) level indicates
additional channel dynamics until the higher-level long-string asymptotics are
actually reached.

For a closed flux tube winding a spatial circle, the quadratic theory has
left- and right-moving oscillators.  Define
\[
  N_L=\sum_{i=1}^{c_\perp}\sum_{n\ge1}n\,N^L_{n,i},
  \qquad
  N_R=\sum_{i=1}^{c_\perp}\sum_{n\ge1}n\,N^R_{n,i}.
\]
Translation along the winding circle gives a quantized longitudinal momentum
\[
  p_\parallel=\frac{2\pi q}{L},
  \qquad q\in\mathbb Z,
\]
and the residual worldsheet translation constraint imposes the level-matching
condition
\begin{equation}
  N_L-N_R=q .
  \label{eq:qcd-string-closed-level-matching}
\end{equation}
At Gaussian order the closed-string energy is
\begin{equation}
  E_{N_L,N_R,q}^{\mathrm{closed}}(L)
  =
  \sigma_kL
  +
  \frac{2\pi}{L}
  \left(
    N_L+N_R-\frac{c_\perp}{12}
  \right)
  +O(L^{-3}),
  \label{eq:qcd-string-closed-gaussian-levels}
\end{equation}
with the \(q\)-dependent kinetic contribution first entering at order
\(L^{-3}\) in the large-\(L\) expansion.  The \(N_L=N_R=0\) case recovers
\eqref{eq:qcd-string-closed-luscher}.

In the Nambu--Goto reference theory these oscillator levels are packaged into
the square-root formulas
\begin{align}
  E_N^{\mathrm{NG,open}}(L)
  &=
  \sigma_R L
  \sqrt{
    1+
    \frac{2\pi}{\sigma_RL^2}
    \left(N-\frac{c_\perp}{24}\right)
  },
  \label{eq:qcd-string-ng-open-levels}\\
  E_{N_L,N_R,q}^{\mathrm{NG,closed}}(L)
  &=
  \sigma_k L
  \sqrt{
    1+
    \frac{4\pi}{\sigma_kL^2}
    \left(N_L+N_R-\frac{c_\perp}{12}\right)
    +
    \left(\frac{2\pi q}{\sigma_kL^2}\right)^2
  },
  \label{eq:qcd-string-ng-closed-levels}
\end{align}
with \eqref{eq:qcd-string-closed-level-matching} imposed in the closed case.
Expanding these formulas gives
\begin{align}
  E_N^{\mathrm{NG,open}}(L)
  &=
  \sigma_RL
  +
  \frac{\pi}{L}
  \left(N-\frac{c_\perp}{24}\right)
  -
  \frac{\pi^2}{2\sigma_RL^3}
  \left(N-\frac{c_\perp}{24}\right)^2
  +O(L^{-5}),
  \label{eq:qcd-string-ng-open-level-expansion}\\
  E_{N_L,N_R,q}^{\mathrm{NG,closed}}(L)
  &=
  \sigma_kL
  +
  \frac{2\pi}{L}
  \left(N_L+N_R-\frac{c_\perp}{12}\right)
  \notag\\
  &\quad
  +
  \frac{2\pi^2}{\sigma_kL^3}
  \left[
    q^2-
    \left(N_L+N_R-\frac{c_\perp}{12}\right)^2
  \right]
  +O(L^{-5}).
  \label{eq:qcd-string-ng-closed-level-expansion}
\end{align}
The square-root formulas are useful because they organize a large set of
levels with the same conventions, but the QCD statement is weaker and more
precise: the Gaussian \(1/L\) structure follows from the assumed gapless
transverse modes, while the \(1/L^3\) and later terms are effective-string
matching data unless a universality theorem for the relevant order and
boundary condition has been established.

\subsection{Oscillator Growth and the Flux-String Hagedorn Scale}

The degeneracy of the oscillator levels in
\eqref{eq:qcd-string-open-oscillator-generating} grows exponentially.  This
growth is a statement about the effective long-string Hilbert space.  The
deconfinement transition of the four-dimensional gauge theory is a property of
the thermal gauge-theory transfer matrix and is discussed in Volume~X.  The
calculation here gives the scale at which a gas of arbitrarily long
effective-string states ceases to have a finite canonical partition function
inside the regime where the long-string description is used.

\paragraph{Colored-partition growth and the string Hagedorn scale.}
Let \(p_c(N)\) be the coefficient of \(q^N\) in
\[
  \prod_{n=1}^\infty(1-q^n)^{-c},
  \qquad c\in\mathbb Z_{>0}.
\]
Then
\begin{equation}
  \log p_c(N)
  =
  2\pi\sqrt{\frac{cN}{6}}+O(\log N)
  \qquad (N\to\infty).
  \label{eq:qcd-string-colored-partition-asymptotic}
\end{equation}
For the Nambu--Goto reference spectrum of a long confining flux string with
\(c_\perp=D-2\) transverse massless fields and tension \(\sigma\), the
oscillator density has the exponential scale
\begin{equation}
  \rho(E)\sim \exp(\beta_{\rm H}^{\rm ws} E),
  \qquad
  \beta_{\rm H}^{\rm ws}
  =
  \sqrt{\frac{\pi c_\perp}{3\sigma}} .
  \label{eq:qcd-string-worldsheet-hagedorn-beta}
\end{equation}
Set \(q=\ee^{-\varepsilon}\), with \(\varepsilon>0\).  The logarithm of the
oscillator generating function is
\[
  -c\sum_{n\ge1}\log(1-\ee^{-n\varepsilon})
  =
  c\sum_{m\ge1}\frac1m\frac{\ee^{-m\varepsilon}}{1-\ee^{-m\varepsilon}} .
\]
As \(\varepsilon\downarrow0\), comparison with the corresponding Riemann sum,
or equivalently the modular transformation of the Dedekind eta function, gives
\begin{equation}
  \log\prod_{n=1}^{\infty}(1-\ee^{-n\varepsilon})^{-c}
  =
  \frac{\pi^2 c}{6\varepsilon}+O(\log\varepsilon).
  \label{eq:qcd-string-partition-small-epsilon}
\end{equation}
Cauchy's coefficient formula writes \(p_c(N)\) as the contour integral of
\(q^{-N-1}\prod_n(1-q^n)^{-c}\).  The exponential part of the integrand on the
positive saddle is
\[
  N\varepsilon+\frac{\pi^2 c}{6\varepsilon}.
\]
The stationary point is
\[
  \varepsilon_*=\pi\sqrt{\frac{c}{6N}},
\]
and the saddle exponent is
\[
  N\varepsilon_*+\frac{\pi^2 c}{6\varepsilon_*}
  =
  2\pi\sqrt{\frac{cN}{6}} .
\]
Gaussian fluctuations and the \(O(\log\varepsilon)\) term in
\eqref{eq:qcd-string-partition-small-epsilon} contribute only logarithmic
corrections to \(\log p_c(N)\), proving
\eqref{eq:qcd-string-colored-partition-asymptotic}.

For an open Nambu--Goto reference string at large excitation number,
\eqref{eq:qcd-string-ng-open-levels} gives
\[
  E^2=2\pi\sigma N+O(N^0)
\]
in the oscillator-dominated regime.  Substituting
\(N=E^2/(2\pi\sigma)+O(1)\) into
\eqref{eq:qcd-string-colored-partition-asymptotic} gives
\[
  \log\rho(E)
  =
  2\pi\sqrt{\frac{c_\perp}{6}\frac{E^2}{2\pi\sigma}}
  +O(\log E)
  =
  E\sqrt{\frac{\pi c_\perp}{3\sigma}}+O(\log E).
\]
For a closed string with level matching at zero longitudinal momentum,
\(N_L=N_R=N\).  The degeneracy is \(p_{c_\perp}(N)^2\), while
\eqref{eq:qcd-string-ng-closed-levels} gives
\[
  E^2=8\pi\sigma N+O(N^0).
\]
The doubled entropy and the doubled energy coefficient combine to the same
\(\beta_{\rm H}^{\rm ws}\).  This proves
\eqref{eq:qcd-string-worldsheet-hagedorn-beta}.

\begin{controlledapproximation}[Flux-tube Hagedorn scale]
\label{ca:qcd-string-hagedorn-interpretation}
Equation~\eqref{eq:qcd-string-worldsheet-hagedorn-beta} is a consequence of
three declared inputs: a stable long flux-tube sector, \(c_\perp\) massless
transverse fields with the oscillator generating function above, and the
Nambu--Goto high-level relation between excitation level and energy.  It is a
worldsheet-density statement.  A four-dimensional deconfinement temperature
requires the thermal gauge-theory path integral, center symmetry and
matter-content data, and the infinite-volume limit.  Equality between a
thermal transition scale and the worldsheet Hagedorn scale is an additional
dynamical statement whose proof would have to relate the thermal transfer
matrix to the long-string density
\eqref{eq:qcd-string-worldsheet-hagedorn-beta}.
\end{controlledapproximation}

\subsection{Static Observables, Lattice Extraction, and the Meaning of
String Tension}

The static potential and the effective string are nonperturbative objects.
Their definition does not use perturbative gluon states.  It uses the Hilbert
space obtained from the gauge theory with external static sources, or
equivalently the transfer matrix acting on gauge-invariant line-operator
states.  On a Euclidean lattice with spacing \(a_{\rm lat}\), a rectangular
Wilson loop is the trace of the ordered product of link variables around the
rectangle.  In the continuum notation used in this chapter it is written
\[
  W_R(C)=\operatorname{tr}_R\,
  \mathcal P\exp\left(i\oint_C A_\mu\,\dd x^\mu\right),
\]
where \(R\) is a representation of the gauge group and
\(\mathcal P\) denotes ordering along the contour.  The continuum expression
is a compact notation for the renormalized line operator obtained from a
chosen regulator; a Wilson loop with cusps or corners has additional local
renormalizations at those singular points.

The area-law statement relevant for the static flux tube is therefore a
statement after local perimeter and corner terms have been separated.  For
large rectangles \(C_{L,T}\), a pure gauge theory in a confining center sector
has
\begin{equation}
  -\log \langle W_R(C_{L,T})\rangle
  =
  \sigma_k LT+\rho_R(L+T)+\kappa_R
  +o(LT)
  \label{eq:qcd-string-rectangular-area-law}
\end{equation}
when both \(L\) and \(T\) are large in a regime where the lowest flux-tube
channel of \(N\)-ality \(k=k_R\) dominates.  The perimeter coefficient
\(\rho_R\) and the corner coefficient \(\kappa_R\) are not the string tension.
They depend on the definition of the line operator and on its ultraviolet
renormalization.  The coefficient \(\sigma_k\) is the asymptotic energy per
unit length of the lowest flux tube in the corresponding center-charge sector.

\paragraph{Rectangular area coefficient and static-potential slope.}
The equality between the rectangular area coefficient and the large-distance
static-potential slope is a transfer-matrix extraction, not an independent
proof of confinement.  Assume the static-source spectral expansion
\eqref{eq:qcd-string-static-source-spectral-expansion}, assume a nonzero
overlap with the lowest state for all sufficiently large \(L\), and assume
that the gap above \(E_0(L)\) is bounded below by a positive function whose
large-\(T\) suppression is uniform in the large-\(L\) window under
consideration.  If
\[
  E_0(L)=\sigma_k L+\mu_R+o(L)
  \qquad (L\to\infty),
\]
then the coefficient of \(LT\) in
\(-\log\langle W_R(C_{L,T})\rangle\), with \(T\to\infty\) taken before the
large-\(L\) slope, is \(\sigma_k\).  Conversely, if the rectangular Wilson
loop has the form \eqref{eq:qcd-string-rectangular-area-law} with a
controlled transfer-matrix limit at each fixed \(L\), then
\[
  \lim_{L\to\infty}\frac{E_0(L)}{L}=\sigma_k .
\]
Indeed, writing the spectral expansion in the form
\[
  \langle W_R(C_{L,T})\rangle
  =
  c_0(L)\ee^{-E_0(L)T}
  \left[
    1+
    \sum_{\alpha>0}\frac{c_\alpha(L)}{c_0(L)}
    \ee^{-(E_\alpha(L)-E_0(L))T}
  \right],
\]
where \(c_0(L)\ne0\) by hypothesis.  The uniform gap assumption implies that
the logarithm of the bracket divided by \(T\) vanishes as \(T\to\infty\), so
\[
  -\lim_{T\to\infty}\frac1T
  \log \langle W_R(C_{L,T})\rangle
  =
  E_0(L).
\]
If \(E_0(L)=\sigma_kL+\mu_R+o(L)\), the coefficient multiplying \(LT\) in the
large-rectangle logarithm is \(\sigma_k\); the \(L\)-independent endpoint
energy \(\mu_R\), the overlap \(c_0(L)\), perimeter terms, and corner terms
do not contribute to the double slope
\(\lim_{L\to\infty}L^{-1}\lim_{T\to\infty}T^{-1}(-\log\langle W\rangle)\).
The converse is the same identity read in the opposite direction: the
transfer-matrix limit extracts \(E_0(L)\), and the assumed rectangular area
coefficient fixes the large-\(L\) slope of this extracted energy.

On a Euclidean time circle of circumference \(\beta\), the corresponding
static observable is the Polyakov loop
\[
  P_R(\mathbf x)=
  \operatorname{tr}_R\,
  \mathcal P\exp\left(i\int_0^\beta A_0(\tau,\mathbf x)\,\dd\tau\right).
\]
The two-source free energy is defined by
\[
  F_R(L;\beta)
  =
  -\frac1\beta
  \log
  \langle P_R(\mathbf 0)^\dagger P_R(\mathbf L)\rangle_\beta .
\]
When the zero-temperature transfer-matrix limit is taken by sending
\(\beta\to\infty\), \(F_R(L;\beta)\) approaches the same static energy
extracted from a long rectangular Wilson loop, up to the vacuum-energy and
line-renormalization conventions used in defining the loop.  At finite
temperature in pure Yang--Mills theory, the expectation value of a single
fundamental Polyakov loop is charged under the temporal center symmetry and is
therefore an order parameter for that finite-temperature symmetry after the
infinite-volume limit is specified.  This is related to, but not identical
with, the zero-temperature string tension \(\sigma_k\).  The latter is a
property of the large-distance static-source spectrum.

In practical lattice calculations one does not measure
\(\sigma_k\) by fitting a bare Wilson loop alone.  One constructs a family of
gauge-invariant operators in the same static-source sector, computes their
Euclidean correlation matrix, and uses a variational transfer-matrix problem
to isolate the lowest and excited energies.  The effective string expansion
then describes the large-\(L\) behavior of these extracted energies.  This is
the operational reason that the QCD string is a nonperturbative observable:
it is a statement about the spectrum selected by gauge-invariant line
operators, not about perturbative gluons propagating between color charges.

\subsection{Baryonic Flux Networks}

Baryonic confinement is not a copy of the mesonic static potential with one
more endpoint.  In an \(SU(N_c)\) gauge theory a gauge-invariant static baryon
operator is made by joining \(N_c\) fundamental Wilson lines to a common color
junction and contracting their color indices with
\(\epsilon_{i_1\cdots i_{N_c}}\).  For paths
\(\Gamma_r\) running from a junction point \(y\) to source positions
\(\mathbf x_r\),
\[
  \mathcal B_{\Gamma_1,\ldots,\Gamma_{N_c}}
  =
  \epsilon_{i_1\cdots i_{N_c}}
  U(\Gamma_1)^{i_1}{}_{j_1}\cdots
  U(\Gamma_{N_c})^{i_{N_c}}{}_{j_{N_c}}
  Q^{j_1}(\mathbf x_1)\cdots Q^{j_{N_c}}(\mathbf x_{N_c}) .
\]
Here \(Q^j(\mathbf x_r)\) denotes a static fundamental source insertion; it
is not a dynamical asymptotic quark state.
The determinant-one property of \(SU(N_c)\) makes the contraction at the
junction gauge invariant.  For \(N_c=3\), and for source separations large
compared with the inverse mass gap, the leading string model predicts an
energy proportional to the length of the minimal three-pronged network, not
to a unique pairwise potential.  Endpoint and junction energies are local
terms and must be separated from the coefficient of the large lengths:
\begin{equation}
  V_{3Q}(\mathbf x_1,\mathbf x_2,\mathbf x_3)
  =
  \sigma\,L_Y(\mathbf x_1,\mathbf x_2,\mathbf x_3)
  +\mu_1+\mu_2+\mu_3+\mu_J+o(L_Y).
  \label{eq:qcd-string-baryon-y-potential}
\end{equation}

\begin{proposition}[Steiner length for a three-source baryonic string]
\label{prop:qcd-string-steiner-y-length}
Let \(a=|\mathbf x_2-\mathbf x_3|\),
\(b=|\mathbf x_3-\mathbf x_1|\), and
\(c=|\mathbf x_1-\mathbf x_2|\) be the side lengths of a Euclidean triangle,
and let \(\Delta\) be its area.  If every angle of the triangle is less than
\(2\pi/3\), the length minimizing
\(\sum_{r=1}^3|\mathbf y-\mathbf x_r|\) is realized by a point
\(\mathbf y_F\) from which the three segments meet at angles \(2\pi/3\), and
\begin{equation}
  L_Y^2
  =
  \frac12
  \left(
    a^2+b^2+c^2+4\sqrt3\,\Delta
  \right).
  \label{eq:qcd-string-steiner-length-formula}
\end{equation}
If one angle is at least \(2\pi/3\), the minimizing network collapses to the
two sides adjacent to that angle.
\end{proposition}

\begin{proof}
For an interior minimizer \(\mathbf y_F\), vary
\(\mathbf y\mapsto \mathbf y+\delta\mathbf y\).  The first variation of
\(\sum_r|\mathbf y-\mathbf x_r|\) is
\[
  \delta\mathbf y\cdot
  \sum_{r=1}^3
  \frac{\mathbf y_F-\mathbf x_r}
       {|\mathbf y_F-\mathbf x_r|}.
\]
It vanishes for all \(\delta\mathbf y\) precisely when the three unit vectors
from the sources to the junction sum to zero.  Three unit vectors sum to zero
only when their pairwise angles are \(2\pi/3\).

Set \(r_i=|\mathbf y_F-\mathbf x_i|\).  Because the angles at the junction are
\(2\pi/3\),
\[
  a^2=r_2^2+r_3^2+r_2r_3,\qquad
  b^2=r_3^2+r_1^2+r_3r_1,\qquad
  c^2=r_1^2+r_2^2+r_1r_2 .
\]
Also the total area is the sum of the three triangles with common vertex
\(\mathbf y_F\):
\[
  \Delta
  =
  \frac{\sqrt3}{4}
  (r_1r_2+r_2r_3+r_3r_1).
\]
Therefore
\[
  \frac12(a^2+b^2+c^2+4\sqrt3\,\Delta)
  =
  r_1^2+r_2^2+r_3^2
  +2(r_1r_2+r_2r_3+r_3r_1)
  =
  (r_1+r_2+r_3)^2 ,
\]
which proves \eqref{eq:qcd-string-steiner-length-formula}.  If an angle is
at least \(2\pi/3\), the stationarity condition cannot be realized by a point
inside the triangle with all three meeting angles \(2\pi/3\).  The convexity
of the sum of distances then places the minimum on the boundary; reducing the
same one-dimensional problem along the boundary gives the vertex at the large
angle, and the network is the union of the two adjacent sides.
\end{proof}

Figure~\ref{fig:qcd-string-baryonic-y-network} illustrates the Steiner-network
length entering the baryonic flux-string model: three strings meet at
\(2\pi/3\) when the Fermat point lies inside the source triangle, and otherwise
the minimum lies on the boundary.

\begin{figure}[H]
  \centering
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \coordinate (A) at (-2.2,-1.0);
    \coordinate (B) at (2.2,-1.0);
    \coordinate (C) at (0.35,1.65);
    \coordinate (F) at (0.05,-0.05);
    \draw[gray!70,thick] (A)--(B)--(C)--cycle;
    \draw[purple!75!black,very thick] (F)--(A);
    \draw[purple!75!black,very thick] (F)--(B);
    \draw[purple!75!black,very thick] (F)--(C);
    \fill[blue!75!black] (A) circle (2.6pt);
    \fill[blue!75!black] (B) circle (2.6pt);
    \fill[blue!75!black] (C) circle (2.6pt);
    \fill[purple!80!black] (F) circle (2.4pt);
    \node[below left] at (A) {\(\mathbf x_1\)};
    \node[below right] at (B) {\(\mathbf x_2\)};
    \node[above] at (C) {\(\mathbf x_3\)};
    \node[right,purple!80!black] at (F) {\(\mathbf y_F\)};
    \draw[->,gray!70] ($(F)+(0.45,-0.15)$) arc[start angle=-18,end angle=-138,radius=0.47];
    \draw[->,gray!70] ($(F)+(-0.25,0.39)$) arc[start angle=122,end angle=2,radius=0.47];
    \node[gray!80!black] at (-0.55,-0.48) {\(2\pi/3\)};
    \node[gray!80!black] at (0.58,0.42) {\(2\pi/3\)};
    \node[purple!75!black,align=center] at (0,-1.65)
      {\(L_Y=|\mathbf y_F-\mathbf x_1|+
      |\mathbf y_F-\mathbf x_2|+|\mathbf y_F-\mathbf x_3|\)};
  \end{tikzpicture}
  \caption{For three static fundamental sources in pure \(SU(3)\) Yang--Mills
  theory, the long-distance baryonic string model uses a gauge-invariant
  color junction.  When all triangle angles are below \(2\pi/3\), the leading
  network is the Steiner tree whose three segments meet at \(2\pi/3\).}
  \label{fig:qcd-string-baryonic-y-network}
\end{figure}

For an equilateral triangle of side \(a\),
\eqref{eq:qcd-string-steiner-length-formula} gives
\[
  L_Y=\sqrt3\,a .
\]
The frequently used pairwise comparison length
\[
  L_\Delta=\frac12(a+b+c)
\]
would give \(L_\Delta=3a/2\) in the same equilateral geometry.  These are
different large-distance hypotheses for the static baryon sector, not two
forms of the same formula.  The gauge-invariant object in the baryonic
Wilson-line construction contains a junction, and the effective string
description must decide, by matching to the gauge theory, whether the
long-distance energy is governed by a junction network, by an approximately
pairwise regime over the available distances, or by a crossover between the
two.

\subsection{Closed Flux Tubes, Glueball Channels, and Large-N String Coupling}

A closed flux tube is cleanly defined on a spatial torus by a spatial
Polyakov loop winding a noncontractible cycle,
\[
  \mathcal P_R(\tau)
  =
  \operatorname{tr}_R\,
  \mathcal P\exp\left(i\oint_{S^1_L}A_i(\tau,\mathbf x)\,\dd x^i\right).
\]
Its Euclidean time correlator has a transfer-matrix expansion
\[
  \langle \mathcal P_R(\tau)^\dagger\mathcal P_R(0)\rangle
  =
  \sum_n
  |\langle n;k,L|\mathcal P_R(0)\vac\rangle|^2
  \ee^{-E_n^{(k)}(L)\tau},
\]
where \(k\) is the center charge of \(R\).  The closed-string formulas
\eqref{eq:qcd-string-closed-luscher},
\eqref{eq:qcd-string-closed-gaussian-levels}, and
\eqref{eq:qcd-string-ng-closed-levels} apply to this winding sector under the
large-\(L\) effective-string hypotheses stated above.

This winding-sector statement is sharper than the phrase ``glueballs are
closed strings.''  A glueball is a state in the vacuum sector created by local
gauge-invariant operators such as
\(\operatorname{tr}(F_{\mu\nu}F^{\mu\nu})\) and
\(\operatorname{tr}(F_{\mu\nu}\widetilde F^{\mu\nu})\).  A noncontractible
winding flux tube belongs to a different finite-volume sector.  A
contractible closed flux loop, by contrast, can shrink to a size of order
\(\sigma^{-1/2}\), where the long-string derivative expansion is no longer
controlled.  Thus the closed-string picture of glueball spectra and Regge
trajectories is a dynamical large-\(N_c\) and effective-string hypothesis, not
a consequence of the Gaussian Luscher calculation.  What is nonperturbatively
well defined are the gauge-invariant correlators and their spectral
decomposition; the string interpretation is an additional infrared
organization whose range must be tested or derived.

Large \(N_c\) supplies the natural weak-coupling parameter for such a string
interpretation once a confining flux-tube spectrum exists.  The ribbon-graph
counting in the preceding section gives a connected contribution of order
\[
  N_c^{2-2h-b}
\]
for a surface of genus \(h\) with \(b\) color-trace boundaries before any
chosen normalization of the external loop operators.  A handle is therefore
suppressed by \(N_c^{-2}\).  Identifying the closed-string genus expansion as
\(g_s^{2h-2}\) gives the parametric relation
\[
  g_s\sim \frac1{N_c}.
\]
The same counting implies that splitting and joining processes of normalized
closed-string states carry the corresponding factors of \(N_c^{-1}\) fixed by
the number of topology-changing vertices.  This statement is a consequence of
color topology; it is not a derivation of the worldsheet action.  The
microscopic input still required is the existence of a
confining large-\(N_c\) gauge theory with isolated flux-tube and glueball
states in the relevant channels.

Three-dimensional confining gauge theories are especially clean testbeds for
the same logic.  In \(D=3\) there is a single transverse displacement field,
so the universal terms become
\[
  \Delta E_0^{\mathrm{open}}(L)=-\frac{\pi}{24L},
  \qquad
  \Delta E_0^{\mathrm{closed}}(L)=-\frac{\pi}{6L}.
\]
The reduced transverse symmetry makes the lattice spectral assignments less
crowded than in four dimensions, while the effective-string expansion has the
same conceptual structure: line operators define the sector, the transfer
matrix extracts energies, the free transverse field gives the universal
Casimir term, and later coefficients encode symmetry constraints and
microscopic matching.

With dynamical fundamental quarks the fundamental one-form symmetry is
explicitly broken.  The static-source Hilbert space then contains states made
from two color-singlet static-light mesons, and the ground-state energy at very
large \(L\) approaches the two-meson threshold rather than growing without
bound.  A flux tube can still be a well-defined long-lived state over a range
of \(L\), and its worldsheet effective theory can describe the corresponding
resonant or metastable channel, but the limiting string tension
\eqref{eq:qcd-string-tension-definition} is no longer an order parameter in the
same sense as in pure Yang--Mills theory.  This distinction is essential when
using string language in QCD: line operators, static-source sectors, and
calorimetric observables are physical, and asymptotic final states are
color-singlet hadrons.

\section{Heavy-Quark Effective Theory and the Wilson-Line Limit}
\label{sec:qcd-hqet-wilson-line-limit}

The static-source sector just described is also the first term in a
controlled expansion for a dynamical quark whose mass is much larger than the
hadronic scale.  This expansion is called heavy-quark effective theory
(HQET).  Its mathematical content is the regulator-dependent comparison of
gauge-invariant QCD matrix elements with gauge-invariant matrix elements of a
local expansion in inverse powers of the heavy mass.  The heavy quark is still
a colored field variable, so physical matrix elements are made gauge invariant
by the same Wilson-line and color-singlet constructions used elsewhere in this
chapter.

Throughout this section the spacetime metric is mostly plus,
\(\eta_{\mu\nu}=\operatorname{diag}(-,+,+,+)\).  A heavy-quark velocity is a
future-directed timelike unit vector \(v^\mu\) satisfying \(v^2=-1\).  The
orthogonal projector onto the three-plane transverse to \(v\) is
\begin{equation}
  \Pi_v{}^\mu{}_\nu=\delta^\mu{}_\nu+v^\mu v_\nu,
  \qquad
  D_\perp^\mu=\Pi_v{}^\mu{}_\nu D^\nu .
  \label{eq:hqet-transverse-projector}
\end{equation}
The identity \(v_\mu D_\perp^\mu=0\) follows immediately from \(v^2=-1\).  A
near-mass-shell heavy momentum is written
\begin{equation}
  p^\mu=m_Qv^\mu+k^\mu ,
  \label{eq:hqet-momentum-split}
\end{equation}
where \(m_Q\) is the heavy-quark mass coordinate in the chosen regulated QCD
Lagrangian and the residual momentum \(k^\mu\) is required to be small
compared with \(m_Q\) on the states or test functions under consideration.
The split is not unique: replacing \(v\) by \(v+\delta v\), with
\(v\cdot\delta v=0\) to first order, and replacing \(k\) by \(k-m_Q\delta v\)
leaves \(p\) fixed.  This elementary redundancy is the kinematic origin of
reparametrization constraints on the effective action.

Let \(\gamma_v=\gamma_\mu v^\mu\).  Since \(\gamma_v^2=-1\), the operator
\(\ii\gamma_v\) has eigenvalues \(\pm1\), and
\begin{equation}
  P_v^\pm=\frac12(1\pm\ii\gamma_v)
  \label{eq:hqet-spin-projectors}
\end{equation}
are complementary spinor projections.  The sign in
\eqref{eq:hqet-spin-projectors} is tied to the phase convention below.  Define
\(\tau_v(x)=-v\cdot x\); in the rest frame \(v=(1,0,0,0)\), this is the
ordinary time coordinate.  A heavy Dirac field \(Q\) is decomposed at fixed
velocity by
\begin{equation}
  Q(x)=\ee^{-\ii m_Q\tau_v(x)}\bigl(h_v(x)+H_v(x)\bigr),
  \qquad
  P_v^+h_v=h_v,\qquad P_v^-H_v=H_v .
  \label{eq:hqet-field-split}
\end{equation}
The phase removes the rest-energy oscillation associated with \(m_Qv^\mu\).
The field \(H_v\) carries the complementary Dirac components and is separated
from \(h_v\) by a gap of order \(m_Q\) in the free theory.

\begin{definition}[HQET regulator specification]
\label{def:hqet-regulator-specification}
A finite-cutoff HQET regulator specification at velocity \(v\) consists of:
\begin{enumerate}
  \item A gauge-covariant ultraviolet regulator for QCD with the heavy field
  \(Q\) and mass coordinate \(m_Q\), together with an infrared domain of states
  or test functions whose residual momenta obey \(|k|/m_Q\le\varepsilon\ll1\).
  \item The projected field \(h_v=P_v^+h_v\), transforming in the fundamental
  representation of the gauge group, and a Wilsonian procedure which
  integrates out \(H_v\) and modes whose virtuality is of order \(m_Q\).
  \item A local gauge-covariant action
  \begin{equation}
    S_{\rm HQET}^{(N)}[h_v,A]
    =
    \int \bar h_v\,\ii v\cdot D\,h_v\,\dd^4x
    +\sum_{r=1}^N\frac1{(2m_Q)^r}
      \sum_a C_{r,a}(\mu,m_Q)\int \mathcal O_{r,a}(x)\,\dd^4x ,
    \label{eq:hqet-local-action}
  \end{equation}
  where each \(\mathcal O_{r,a}\) is a local gauge-invariant density built
  from \(h_v,\bar h_v,F_{\mu\nu}\), and transverse covariant derivatives
  \(D_\perp\), modulo equations of motion and integration by parts in the
  chosen coordinate system on local functionals.
  \item Renormalized composite operators, for example
  \[
    \bar q\,\Gamma Q
    \longmapsto
    \sum_{r,a}\frac1{(2m_Q)^r}C_{\Gamma,r,a}(\mu,m_Q)\,
    \mathcal J_{\Gamma,r,a}^{\rm HQET}[q,h_v,A],
  \]
  with the same regulator and subtraction data.  A current matching statement
  is a statement about matrix elements of these renormalized operators, not an
  identity between classical monomials.
\end{enumerate}
The leading action is normalized so that the coefficient of
\(\bar h_v\ii v\cdot D h_v\) is one.  The first dimension-five operators may
be represented by
\begin{equation}
  \mathcal O_{\rm kin}=\bar h_vD_{\perp\mu}D_\perp^\mu h_v,
  \qquad
  \mathcal O_{\rm mag}=\bar h_v\sigma_{\mu\nu}F^{\mu\nu}h_v ,
  \label{eq:hqet-dimension-five-operators}
\end{equation}
with coefficients fixed by the chosen matching convention.  In a
reparametrization-compatible scheme the coefficient of the kinetic operator is
fixed relative to the leading term; the chromomagnetic coefficient is a
genuine short-distance matching coefficient.
\end{definition}

The leading HQET equation is the equation for parallel transport along the
heavy-quark worldline.  This is the precise sense in which an infinitely heavy
quark becomes a Wilson line.

\paragraph{Static propagator as a Wilson line.}
Let \(x(s)=x_0+sv\), and let
\begin{equation}
  W_v(s_2,s_1)=\mathcal P\exp\left\{
    \ii\int_{s_1}^{s_2}v^\mu A_\mu^a(x(s))t_\square^a\,\dd s
  \right\}
  \label{eq:hqet-worldline-wilson-line}
\end{equation}
be the fundamental Wilson transporter along the heavy trajectory.  At finite
regulator and with background \(A_\mu\) smooth on the scale of the regulator,
the retarded Green kernel of the leading HQET operator on this trajectory is
\begin{equation}
  G_v(s_2,s_1)=\theta(s_2-s_1)\,W_v(s_2,s_1)\,P_v^+ ,
  \label{eq:hqet-static-green-kernel}
\end{equation}
up to the conventional overall factor fixed by the normalization of the
first-order inverse.  In particular, every gauge-invariant heavy-light or
heavy-heavy correlator at leading power contains the Wilson line carried by
the heavy source.
The calculation is the Green-kernel inversion of the first-order covariant
derivative.  The transporter satisfies the initial-value problem
\[
  \frac{\dd}{\dd s_2}W_v(s_2,s_1)
  =
  \ii v^\mu A_\mu^a(x(s_2))t_\square^a W_v(s_2,s_1),
  \qquad W_v(s_1,s_1)=\mathbf 1 .
\]
Equivalently \(v\cdot D_{x(s_2)}W_v(s_2,s_1)=0\), using the chapter
convention \(D_\mu=\partial_\mu-\ii A_\mu\).  Applying the first-order operator
to \(\theta(s_2-s_1)W_v(s_2,s_1)P_v^+\) gives only the derivative of the step
function, because the covariant derivative annihilates the transporter for
\(s_2\ne s_1\).  This is the delta distribution on the oriented line.  The
projector \(P_v^+\) records the spinor components retained in the heavy field.
Gauge covariance follows from the same Wilson-transporter endpoint
transformation law displayed in
\eqref{eq:qcd-wilson-transporter-gauge-transformation}.

Consider \(N_h\) heavy flavors whose masses have been removed by separate
phases and whose residual dynamics are described by the leading HQET action at
the same velocity \(v\).  The leading action
\[
  S_0=\sum_{A=1}^{N_h}\int \bar h_{v,A}\,\ii v\cdot D\,h_{v,A}\,\dd^4x
\]
is invariant under unitary rotations of the heavy-flavor labels and under
unitary rotations of the projected spin components commuting with \(P_v^+\).
These symmetries act on the heavy variables alone and leave the gauge and
light-quark fields fixed.
The operator \(\ii v\cdot D\) is scalar in the heavy-flavor labels and
contains no gamma matrix acting inside the projected spin space.  If
\(h_{v,A}\mapsto U_A{}^B Sh_{v,B}\), where \(U\) is a unitary matrix on flavor
labels and \(S\) is unitary on the image of \(P_v^+\), then
\(\bar h_{v,A}\ii v\cdot D h_{v,A}\) is unchanged after summing over \(A\).
The Wilson-line form of the propagator makes the same statement in correlator
language: at leading power the heavy source transports color, while its spin
and heavy-flavor labels are parallel spectators.

For a free heavy particle with \(p^\mu=m_Qv^\mu+k^\mu\) and \(p^2+m_Q^2=0\),
the residual momentum obeys
\begin{equation}
  2m_Q\,v\cdot k+k^2=0 .
  \label{eq:hqet-onshell-residual}
\end{equation}
In the rest frame, writing \(k^\mu=(E_{\rm res},\mathbf k)\), this gives
\begin{equation}
  E_{\rm res}
  =
  \sqrt{m_Q^2+|\mathbf k|^2}-m_Q
  =
  \frac{|\mathbf k|^2}{2m_Q}
  -\frac{|\mathbf k|^4}{8m_Q^3}
  +O\!\left(\frac{|\mathbf k|^6}{m_Q^5}\right).
  \label{eq:hqet-free-dispersion}
\end{equation}
Thus the first correction to the static theory is the transverse kinetic
energy.  Gauge covariance promotes the transverse momentum to \(D_\perp\),
giving the operator \(\mathcal O_{\rm kin}\) in
\eqref{eq:hqet-dimension-five-operators}.
Indeed, substituting \(p=m_Qv+k\) and \(v^2=-1\) into \(p^2+m_Q^2=0\) gives
\eqref{eq:hqet-onshell-residual}.  In the rest frame,
\(p=(m_Q+E_{\rm res},\mathbf k)\), so the positive-energy solution is
\(m_Q+E_{\rm res}=\sqrt{m_Q^2+|\mathbf k|^2}\).  Expanding the square root in
\(|\mathbf k|^2/m_Q^2\) gives \eqref{eq:hqet-free-dispersion}.  The covariant
operator statement follows from replacing the residual transverse momentum by
the transverse covariant derivative in the local Wilsonian expansion; the
coefficient is fixed by matching the free dispersion and, in an interacting
reparametrization-compatible scheme, by the invariance of the momentum split
\eqref{eq:hqet-momentum-split}.

\begin{controlledapproximation}[Meaning of the heavy-mass expansion]
\label{ca:hqet-heavy-mass-expansion}
Fix a regulator, a velocity \(v\), a renormalization scale \(\mu\ll m_Q\), and
a class of gauge-invariant external states and operator insertions whose
residual momenta and invariant hadronic scales are bounded by
\(\Lambda_{\rm had}\).  A truncation of HQET at order \(N\) asserts an
estimate of the form
\[
  \left|
  \langle \mathcal O\rangle_{\rm QCD}
  -
  \langle \mathcal O^{(N)}_{\rm HQET}\rangle_{\rm HQET}
  \right|
  \le
  C_{\mathcal O,N}
  \left(\frac{\Lambda_{\rm had}}{m_Q}\right)^{N+1}
\]
for the chosen smeared matrix elements, after the QCD operator and the HQET
operator have been matched in the same renormalization scheme.  The constant
\(C_{\mathcal O,N}\), the topology in which the estimate is meant, and the
  range of \(\mu\) are part of the assertion.  The Wilson-line limit derived
above supplies the leading operator structure; the existence of a uniform
heavy-mass expansion for a specified family of hadronic matrix elements is
additional dynamical and renormalization information.
\end{controlledapproximation}

\section{Heavy-Light Currents and Heavy-Quark Symmetry}
\label{sec:qcd-hqet-heavy-light-currents}

The static Wilson-line limit becomes predictive only after composite currents
and hadron states are matched.  Let \(q\) be a light quark field and let
\(Q\) be a heavy quark field.  The mostly-plus spinor convention of
Definition~\ref{def:four-dimensional-mostly-plus-spinor-convention} fixes the
Hermitian vector and axial currents with an explicit factor of \(\ii\).  Thus
in this section
\[
  A^\mu_{qQ}=\ii[\bar q\gamma^\mu\gamma_5Q]_{\rm R},
  \qquad
  V^\mu_{qQ}=\ii[\bar q\gamma^\mu Q]_{\rm R}.
\]
More generally a renormalized heavy-light QCD current is a renormalized
gauge-invariant local operator
\[
  J_\Gamma^{\rm QCD}(x)=[\kappa_\Gamma\bar q(x)\Gamma Q(x)]_{\rm R},
\]
where \(\Gamma\) is a fixed spinor matrix, \(\kappa_\Gamma\in\{1,\ii\}\) is
part of the current convention, and the subscript records the regulator and
subtraction convention.  In HQET this current is represented by an expansion
\begin{equation}
  J_\Gamma^{\rm QCD}
  =
  C_\Gamma(m_Q,\mu)\,[\kappa_\Gamma\bar q\Gamma h_v]_{\rm R}
  +
  \sum_a\frac{B_{\Gamma,a}(m_Q,\mu)}{2m_Q}\,
  \mathcal J_{\Gamma,a}^{(1)}
  +O(m_Q^{-2})
  \label{eq:hqet-current-matching}
\end{equation}
inside the chosen class of matrix elements.  The operators
\(\mathcal J_{\Gamma,a}^{(1)}\) are local gauge-invariant HQET currents with
one additional transverse derivative, chromomagnetic insertion, or light-quark
mass insertion, modulo the leading equations of motion in the chosen operator
basis.  Equation \eqref{eq:hqet-current-matching} is a matching statement for
renormalized operator-valued distributions tested on low-residual-momentum
states.

Assume that for a heavy flavor \(Q\) there is an isolated stable pseudoscalar
meson \(P_Q\) and an isolated stable vector meson \(V_Q\) in the same
light-cloud channel.  Their decay constants are defined by
\begin{align}
  \langle 0|\,A^\mu_{qQ}\,|P_Q(p)\rangle
  &=
  f_{P_Q}\,p^\mu,
  \label{eq:hqet-pseudoscalar-decay-constant}\\
  \langle 0|\,V^\mu_{qQ}\,|V_Q(p,\varepsilon)\rangle
  &=
  f_{V_Q}\,M_{V_Q}\,\varepsilon^\mu,
  \label{eq:hqet-vector-decay-constant}
\end{align}
where \(p^2=-M^2\), \(p^0>0\), and the vector polarization satisfies
\(p\cdot\varepsilon=0\).  Momentum eigenstates in QCD are normalized by
\[
  \langle H(p')|H(p)\rangle
  =
  2p^0(2\pi)^3\delta^3(\mathbf p-\mathbf p').
\]
The corresponding HQET residual-momentum state is normalized by
\[
  {}_{\rm HQET}\langle H_v(k')|H_v(k)\rangle_{\rm HQET}
  =
  2v^0(2\pi)^3\delta^3(\mathbf k-\mathbf k'),
\]
with \(p^\mu=M_Hv^\mu+k^\mu\), and the leading state matching is
\begin{equation}
  |H_Q(p=M_Hv+k)\rangle_{\rm QCD}
  =
  \sqrt{M_H}\,|H_v(k)\rangle_{\rm HQET}
  +O(\Lambda_{\rm had}/m_Q)
  \label{eq:hqet-state-normalization-matching}
\end{equation}
for matrix elements in the stated residual-momentum domain.

\paragraph{Heavy-light decay-constant scaling.}
Let
\[
  {}_{\rm HQET}\langle0|\,\ii[\bar q\gamma^\mu\gamma_5h_v]_{\rm R}
  |P_v(0)\rangle_{\rm HQET}
  =
  F_P(\mu)\,v^\mu .
\]
Then the leading HQET matching of the axial current gives
\begin{equation}
  f_{P_Q}\sqrt{M_{P_Q}}
  =
  C_A(m_Q,\mu)F_P(\mu)
  +O(\Lambda_{\rm had}/m_Q).
  \label{eq:hqet-decay-constant-scaling}
\end{equation}
The analogous vector-meson relation has the same static reduced matrix
element up to the spin rotation dictated by the leading static symmetry
described above.
At \(p^\mu=M_{P_Q}v^\mu\), the QCD definition
\eqref{eq:hqet-pseudoscalar-decay-constant} gives
\[
  \langle0|A^\mu_{qQ}|P_Q(p)\rangle_{\rm QCD}
  =
  f_{P_Q}M_{P_Q}v^\mu .
\]
Using the leading current matching
\[
  A^\mu_{qQ}
  =
  C_A(m_Q,\mu)\,\ii[\bar q\gamma^\mu\gamma_5h_v]_{\rm R}
  +O(m_Q^{-1})
\]
and the state relation \eqref{eq:hqet-state-normalization-matching} gives
\[
  \langle0|A^\mu_{qQ}|P_Q(p)\rangle_{\rm QCD}
  =
  C_A\sqrt{M_{P_Q}}\,
  F_P(\mu)v^\mu
  +O(\Lambda_{\rm had}/m_Q).
\]
Equating the coefficients of \(v^\mu\) gives
\eqref{eq:hqet-decay-constant-scaling}.  The vector statement follows because
the leading static action transports color but contains no operator acting on
the projected heavy spin; pseudoscalar and vector states in the same light
cloud are related by the heavy-spin symmetry of the leading theory.

Heavy-to-heavy currents compare two static sectors with velocities \(v\) and
\(v'\).  Define the recoil variable
\begin{equation}
  w=-v\cdot v'\ge1 .
  \label{eq:hqet-recoil-variable}
\end{equation}
For pseudoscalar ground-state mesons in the same light-cloud channel, the
leading HQET vector-current matrix element has the form
\begin{equation}
  {}_{\rm HQET}\langle H_{v'}|\,
  \mathcal V_{Q'Q}^\mu(v',v)\,|H_v\rangle_{\rm HQET}
  =
  \xi(w)(v+v')^\mu ,
  \label{eq:hqet-isgur-wise-definition}
\end{equation}
where
\[
  \mathcal V_{Q'Q}^\mu(v',v)
  =
  \ii[\bar h_{v'}^{(Q')}\gamma^\mu h_v^{(Q)}]_{\rm R}
\]
with the coefficient fixed by coupling the leading HQET action to a background
heavy-flavor gauge field.  The scalar function \(\xi\) is the leading
heavy-quark symmetry form factor.  For the full QCD states this corresponds to
\begin{equation}
  \begin{aligned}
  &\langle H_{Q'}(M_{Q'}v')|\,\ii[\bar Q'\gamma^\mu Q]_{\rm R}\,
  |H_Q(M_Qv)\rangle
  \\
  &\qquad =
  \sqrt{M_{Q'}M_Q}\,
  C_{Q'Q}(m_{Q'},m_Q,\mu)\,
  \xi(w)(v+v')^\mu
  +O(\Lambda_{\rm had}/m_Q,\Lambda_{\rm had}/m_{Q'}).
  \end{aligned}
  \label{eq:hqet-heavy-to-heavy-qcd-form-factor}
\end{equation}
More general spin states are obtained from the same light-cloud overlap by
inserting the appropriate heavy-spin trace factors.

\begin{lemma}[Zero-recoil normalization]
\label{lem:hqet-isgur-wise-normalization}
At zero recoil, \(v'=v\), the leading heavy-quark symmetry form factor obeys
\begin{equation}
  \xi(1)=1
  \label{eq:hqet-isgur-wise-normalization}
\end{equation}
when the current is normalized as the generator of the heavy-flavor symmetry
of the leading HQET action.
\end{lemma}

\begin{proof}
At \(v'=v\), the leading HQET action has a unitary flavor symmetry rotating
the heavy labels \(Q\) and \(Q'\) while leaving the light fields and gauge
field fixed.  This statement is made inside the leading static theory, before
the short-distance matching coefficient \(C_{Q'Q}\) and before \(1/m_Q\)
corrections are restored.  Couple this symmetry to a smooth background
matrix-valued
connection \(B_\mu{}^A{}_B\) by replacing
\(D_\mu h_{v,A}\) with
\((D_\mu\delta_A{}^B-\ii B_\mu{}^B{}_A)h_{v,B}\).  Differentiation of the
action with respect to \(B_\mu{}^Q{}_{Q'}\) defines the current
\(\mathcal V^\mu_{Q'Q}(v,v)\), and the associated charge
\[
  Q_{Q'Q}
  =
  \int_{\Sigma_t}\dd^3x\,n_\mu\,\mathcal V^\mu_{Q'Q}(v,v)
\]
is the generator of the heavy-flavor rotation on the one-heavy-particle
sector.  Here \(n_\mu\) is the future unit normal covector to the quantization
slice.  Choose the isolated ground states in the same light-cloud channel with
the same Hilbert-space norm and with phases fixed so that the flavor
rotation maps one state to the other.  The Ward identity for the background
flavor source then gives
\[
  Q_{Q'Q}|H_v^{(Q)}\rangle=|H_v^{(Q')}\rangle .
\]
Translation covariance in the fixed-velocity sector and the absence of any
other vector built from a single velocity imply that the local current matrix
element has the form \(A\,v^\mu\).  Integrating \(n_\mu A v^\mu\) over the
normalization volume and using the preceding charge action fixes \(A=2\) in
the velocity-state normalization used in \eqref{eq:hqet-isgur-wise-definition}.
Equivalently,
\[
  {}_{\rm HQET}\langle H_v^{(Q')}|
  \mathcal V^\mu_{Q'Q}(v,v)
  |H_v^{(Q)}\rangle_{\rm HQET}
  =
  2v^\mu .
\]
Comparing with \eqref{eq:hqet-isgur-wise-definition} at \(v'=v\) gives
\(\xi(1)=1\).
\end{proof}

\begin{controlledapproximation}[Heavy-light current expansion]
\label{ca:hqet-heavy-light-current-expansion}
A heavy-light or heavy-to-heavy form-factor prediction must specify the
renormalized QCD current, the HQET current basis, the state normalization, the
matching coefficients, and the spectral assumption that the meson states used
in the matrix element are isolated in the relevant channel.  The leading
relations \eqref{eq:hqet-decay-constant-scaling} and
\eqref{eq:hqet-isgur-wise-definition} are symmetry consequences of the static
theory after these data are fixed.  Corrections of order \(1/m_Q\) are matrix
elements of the subleading current operators and of insertions from the
subleading HQET action; their absence or suppression at special kinematic
points is an additional theorem only when the corresponding operator basis and
symmetry argument are displayed.
\end{controlledapproximation}

\section{Nonrelativistic Heavy Pairs and Potential Matching}
\label{sec:qcd-nrqcd-potential-matching}

Quarkonium requires a heavy-mass expansion with two heavy fields and a
specified time foliation.  Choose the rest frame of the heavy pair and write
\(t\) for its time coordinate.  A nonrelativistic heavy quark is represented by
a two-component Pauli field \(\psi(t,\mathbf x)\) in the fundamental
representation.  The corresponding antiquark is represented, after charge
conjugation, by another Pauli field \(\chi(t,\mathbf x)\) which also
transforms in the fundamental representation; its adjoint \(\chi^\dagger\)
creates an antiquark and transforms as a row antifundamental.  With
\[
  D_t=\partial_t-\ii A_0,\qquad
  \mathbf D=\nabla-\ii\mathbf A ,
\]
the leading nonrelativistic action is
\begin{equation}
  S_{\rm NR}^{(0)}
  =
  \int \dd t\,\dd^3\mathbf x\,
  \left[
    \psi^\dagger\left(\ii D_t+\frac{\mathbf D^2}{2m_Q}\right)\psi
    +
    \chi^\dagger\left(\ii D_t+\frac{\mathbf D^2}{2m_Q}\right)\chi
  \right].
  \label{eq:nrqcd-leading-action}
\end{equation}
The sign of the spatial Laplacian in \eqref{eq:nrqcd-leading-action} is fixed
by the convention \(D_i\ee^{\ii\mathbf p\cdot\mathbf x}=\ii p_i
\ee^{\ii\mathbf p\cdot\mathbf x}\) in the free theory.

\begin{definition}[NRQCD regulator and operator specification]
\label{def:nrqcd-regulator-operator-specification}
An NRQCD regulator and operator specification for a heavy flavor of mass
\(m_Q\) consists of:
\begin{enumerate}
  \item A gauge-covariant regulator for QCD and a matching scale below
  \(m_Q\), together with a class of states whose heavy-quark three-momenta
  obey \(|\mathbf p|/m_Q\ll1\).
  \item Pauli fields \(\psi,\chi\), the gauge fields, and a local action of
  the form
  \[
    S_{\rm NRQCD}^{(N)}
    =
    S_{\rm NR}^{(0)}
    +
    \sum_{a}\frac{c_a(\mu,m_Q)}{m_Q^{d_a-4}}
    \int \mathcal O_a^{\rm NR}(t,\mathbf x)\,\dd t\,\dd^3\mathbf x ,
  \]
  where \(\mathcal O_a^{\rm NR}\) ranges over gauge-invariant local operators
  built from \(\psi,\chi,F_{\mu\nu}\), covariant derivatives, Pauli matrices,
  and local four-heavy-field annihilation operators, ordered by the chosen
  velocity and \(1/m_Q\) power counting.
  \item Renormalized heavy-pair composite operators.  A quarkonium interpolator
  or current is part of the specification only after its Wilson-line dressing,
  regulator, and subtraction convention have been specified.
\end{enumerate}
The Pauli magnetic term
\[
  \frac{c_F}{2m_Q}\,
  \psi^\dagger\sigma^iB_i\psi
  -
  \frac{c_F}{2m_Q}\,
  \chi^\dagger\sigma^iB_i\chi,
  \qquad
  B_i=\frac12\epsilon_{ijk}F_{jk},
\]
is the first spin-dependent single-particle correction in a parity-invariant
coordinate system.  Its coefficient is a matching coefficient; it is not
deduced from the leading Schrödinger action alone.
\end{definition}

\paragraph{Free NRQCD dispersion.}
The free equation following from the leading quark term in
\eqref{eq:nrqcd-leading-action} has plane-wave solutions
\[
  \psi(t,\mathbf x)=u\,\ee^{-\ii Et+\ii\mathbf p\cdot\mathbf x}
\]
with
\[
  E=\frac{|\mathbf p|^2}{2m_Q}.
\]
Thus \eqref{eq:nrqcd-leading-action} realizes the leading term of the
relativistic residual-energy expansion
\eqref{eq:hqet-free-dispersion}.
Substitution of the plane wave gives
\[
  \left(E-\frac{|\mathbf p|^2}{2m_Q}\right)u=0,
\]
because \(\nabla^2\ee^{\ii\mathbf p\cdot\mathbf x}
=-|\mathbf p|^2\ee^{\ii\mathbf p\cdot\mathbf x}\).  A nonzero spinor
amplitude \(u\) therefore requires the displayed dispersion relation.

Gauge-invariant quarkonium operators are bilocal at nonzero separation.  Let
\[
  \mathbf x_1=\mathbf R+\frac{\mathbf r}{2},
  \qquad
  \mathbf x_2=\mathbf R-\frac{\mathbf r}{2},
\]
and let \(W(\mathbf x_2,\mathbf x_1;t)\) be the equal-time fundamental Wilson
transporter from \(\mathbf x_1\) to \(\mathbf x_2\).  Define the singlet
bilocal
\begin{equation}
  S(t,\mathbf R,\mathbf r)
  =
  \frac1{\sqrt{N_c}}\,
  \chi^\dagger(t,\mathbf x_2)
  W(\mathbf x_2,\mathbf x_1;t)
  \psi(t,\mathbf x_1).
  \label{eq:pnrqcd-singlet-bilocal}
\end{equation}
An octet bilocal with a color label based at \(\mathbf R\) is represented by
\begin{equation}
  O^a(t,\mathbf R,\mathbf r)
  =
  \chi^\dagger(t,\mathbf x_2)
  W(\mathbf x_2,\mathbf R;t)\,
  t^a\,
  W(\mathbf R,\mathbf x_1;t)
  \psi(t,\mathbf x_1).
  \label{eq:pnrqcd-octet-bilocal}
\end{equation}
The octet transforms in the adjoint representation at \(\mathbf R\); the
singlet is gauge invariant.

\paragraph{Gauge covariance of the bilocals.}
Under a gauge transformation \(U(t,\mathbf x)\),
\[
  \psi(\mathbf x_1)\mapsto U(\mathbf x_1)\psi(\mathbf x_1),
  \qquad
  \chi^\dagger(\mathbf x_2)\mapsto
  \chi^\dagger(\mathbf x_2)U(\mathbf x_2)^{-1},
\]
and
\[
  W(\mathbf x_2,\mathbf x_1)
  \mapsto
  U(\mathbf x_2)W(\mathbf x_2,\mathbf x_1)U(\mathbf x_1)^{-1}.
\]
The endpoint matrices cancel in \eqref{eq:pnrqcd-singlet-bilocal}.  In
\eqref{eq:pnrqcd-octet-bilocal} the endpoint matrices again cancel, while the
generator at \(\mathbf R\) transforms by conjugation,
\(t^a\mapsto U(\mathbf R)t^aU(\mathbf R)^{-1}\), which is the adjoint action.

Potential NRQCD is obtained by integrating out the relative-momentum scale
\(m_Qv_{\rm rel}\) while retaining the lower energy scale
\(m_Qv_{\rm rel}^2\).  The singlet part of the resulting bilocal action has
the form
\begin{equation}
  S_{\rm pNRQCD}^{S}
  =
  \int \dd t\,\dd^3\mathbf R\,\dd^3\mathbf r\,
  S^\dagger
  \left[
    \ii\partial_t
    +\frac{\nabla_{\mathbf r}^2}{m_Q}
    -V_S(r;\mu)
    +\cdots
  \right]S ,
  \label{eq:pnrqcd-singlet-action}
\end{equation}
where \(r=|\mathbf r|\).  The potential \(V_S(r;\mu)\) is a matching
coefficient for the bilocal singlet operator.  In a weak-coupling matching
regime, the tree-level Coulomb exchange gives
\begin{equation}
  V_S^{(0)}(r)=-\frac{g^2C_F}{4\pi r},
  \label{eq:pnrqcd-tree-coulomb-potential}
\end{equation}
with \(C_F=(N_c^2-1)/N_c\) and the same trace-delta coupling used throughout
this chapter.  This formula is invariant under conversion to the common
half-trace convention because \(g_{\rm ht}^2=2g^2\) and
\(C_F^{\rm ht}=C_F/2\).

\paragraph{Tree-level singlet color factor.}
In the trace convention of \eqref{eq:qcd-casimir-definitions}, the color
factor for one-gluon exchange in a normalized quark-antiquark singlet is
\(-C_F\).  Hence the attractive Coulomb potential is
\eqref{eq:pnrqcd-tree-coulomb-potential}.

The normalized singlet color tensor is
\(\delta^i{}_j/\sqrt{N_c}\).  The antiquark generator is the negative
transpose of the quark generator.  Acting with \(t^a\otimes(-t^{aT})\) on the
singlet gives
\[
  -\frac1{\sqrt{N_c}}(t^at^a)^i{}_j
  =
  -C_F\frac{\delta^i{}_j}{\sqrt{N_c}} .
\]
The spatial Fourier transform of the Coulomb kernel is
\[
  \int\frac{\dd^3\mathbf q}{(2\pi)^3}\,
  \frac{\ee^{\ii\mathbf q\cdot\mathbf r}}{|\mathbf q|^2}
  =
  \frac1{4\pi r},
\]
so the product of the exchange kernel \(g^2/|\mathbf q|^2\) with the singlet
color eigenvalue gives \eqref{eq:pnrqcd-tree-coulomb-potential}.

\begin{controlledapproximation}[Ultrasoft dipole sector in weak-coupling pNRQCD]
\label{ca:qcd-pnrqcd-ultrasoft-dipole-window}
A weak-coupling pNRQCD prediction is not specified by the static singlet
potential alone.  After the soft scale \(m_Qv_{\rm rel}\) has been integrated
out, the lower-energy gauge field still fluctuates on the ultrasoft scale
\(m_Qv_{\rm rel}^2\).  The multipole expansion around the center coordinate
\(\mathbf R\) is a separate expansion in
\[
  r\,\partial_{\rm us}\sim v_{\rm rel},
  \qquad r\sim (m_Qv_{\rm rel})^{-1}.
\]
To leading dipole order the singlet and adjoint-octet bilocals are coupled by
the chromoelectric field at \(\mathbf R\),
\begin{equation}
  S_{\rm us}^{SO}
  =
  \int \dd t\,\dd^3\mathbf R\,\dd^3\mathbf r\,
  V_A(r;\mu)\,
  r^i E_i^a(t,\mathbf R)
  \left[O^{a\dagger}(t,\mathbf R,\mathbf r)S(t,\mathbf R,\mathbf r)
  +S^\dagger(t,\mathbf R,\mathbf r)O^a(t,\mathbf R,\mathbf r)\right],
  \label{eq:qcd-pnrqcd-ultrasoft-dipole-coupling}
\end{equation}
where \(V_A\) includes the normalization convention for the gauge field and
for the octet basis used in \eqref{eq:pnrqcd-octet-bilocal}.  In a canonical
half-trace field normalization the same term is usually written with an
explicit gauge coupling; here the normalization is fixed by matching
renormalized QCD matrix elements in the same trace-delta convention as
\eqref{eq:pnrqcd-tree-coulomb-potential}.

The octet is not an optional decoration.  At tree level its color factor is
\(+1/N_c\) in the trace-delta convention, so
\begin{equation}
  V_O^{(0)}(r)=+\frac{g^2}{4\pi N_c r},
  \qquad
  V_O^{(0)}(r)-V_S^{(0)}(r)
  =
  \frac{g^2N_c}{4\pi r}.
  \label{eq:qcd-pnrqcd-tree-octet-gap}
\end{equation}
Thus an ultrasoft chromoelectric fluctuation can virtually promote a singlet
quarkonium state to an octet-heavy-pair state, propagate with the octet
Hamiltonian, and return to the singlet.  For an isolated singlet level
\(|n\rangle\) with singlet Hamiltonian \(H_S|n\rangle=E_n|n\rangle\), the
leading dipole contribution has the renormalized coordinate form
\begin{equation}
  \Delta E_{n,{\rm us}}
  =
  -\ii
  \int_0^\infty \dd t\,
  \langle n|
    r^i\,
    \ee^{-\ii(H_O-E_n)t}\,
    r^j
  |n\rangle\,
  V_A^2\,
  K_{EE}^{ij}(t;\mu)
  +\Delta E_{n,{\rm ct}}(\mu),
  \label{eq:qcd-pnrqcd-ultrasoft-energy-shift}
\end{equation}
with
\[
  K_{EE}^{ij}(t;\mu)
  =
  \left\langle
  E_i^a(t,\mathbf R)\,
  \Phi_{\rm adj}^{ab}(t,0;\mathbf R)\,
  E_j^b(0,\mathbf R)
  \right\rangle_{\rm us,R,\mu}.
\]
The adjoint Wilson transporter \(\Phi_{\rm adj}\) is required for gauge
covariance of the chromoelectric correlator.  The counterterm
\(\Delta E_{n,{\rm ct}}\) subtracts the ultraviolet part already assigned to
the matched potentials and fixes the \(\mu\)-dependence of \(V_S\).  If the
octet-ultrasoft intermediate channel crosses a physical cut, the same
resolvent gives an imaginary part and hence a transition width; below such
cuts it gives a Lamb-shift-like real contribution.  Either way, the physical
object is an energy or width extracted after integrating over ultrasoft
fluctuations, not the static source geometry alone.

A controlled use of \eqref{eq:qcd-pnrqcd-ultrasoft-energy-shift} must state:
the singlet and octet Hamiltonians and mass scheme; the matching scale and
coefficient \(V_A\); the renormalized chromoelectric correlator and its
counterterm convention; the spectral window separating the chosen singlet
level from nearby thresholds; and residual estimates for omitted multipoles,
higher potentials, relativistic corrections, and nonperturbative ultrasoft
input.  Without this data, a potential-model spectrum is only the leading
coordinate of the pNRQCD expansion.
\end{controlledapproximation}

\paragraph{Leading Coulombic \(1S\) quarkonium level.}
Assume a weak-coupling pNRQCD chart in which the spin-independent singlet
Hamiltonian is truncated to
\begin{equation}
  H_C
  =
  -\frac{\nabla_{\mathbf r}^2}{m_Q}
  -\frac{\alpha_C}{r},
  \qquad
  \alpha_C:=\frac{g^2C_F}{4\pi},
  \qquad r=|\mathbf r|.
  \label{eq:qcd-coulombic-pnrqcd-hamiltonian}
\end{equation}
Here \(m_Q\) is the heavy-quark mass coordinate in the same short-distance
scheme as the potential coordinate.  The normalized \(1S\) eigenfunction
\begin{equation}
  \psi_{1S}(\mathbf r)
  =
  \left(\frac{\kappa^3}{\pi}\right)^{1/2}\ee^{-\kappa r},
  \qquad
  \kappa:=\frac{m_Q\alpha_C}{2},
  \label{eq:qcd-coulombic-one-s-wavefunction}
\end{equation}
has energy
\begin{equation}
  E_{1S}^{(0)}
  =
  -\frac{m_Q\alpha_C^2}{4},
  \qquad
  |\psi_{1S}(0)|^2=
  \frac{(m_Q\alpha_C)^3}{8\pi}.
  \label{eq:qcd-coulombic-one-s-energy-origin}
\end{equation}
The corresponding leading spin-averaged quarkonium mass coordinate is
\[
  M_{1S}^{(0)}=2m_Q+E_{1S}^{(0)}
\]
before the additive mass--potential rearrangements of
Section~\ref{sec:qcd-heavy-mass-static-energy} and before ultrasoft,
relativistic, and spin-dependent corrections are included.

For \(r>0\),
\[
  \nabla^2\ee^{-\kappa r}
  =
  \left(\kappa^2-\frac{2\kappa}{r}\right)\ee^{-\kappa r}.
\]
Thus
\[
  H_C\ee^{-\kappa r}
  =
  \left[
    -\frac{\kappa^2}{m_Q}
    +
    \left(\frac{2\kappa}{m_Q}-\alpha_C\right)\frac1r
  \right]\ee^{-\kappa r}.
\]
With \(\kappa=m_Q\alpha_C/2\) the \(1/r\) term cancels and the eigenvalue is
\(-\kappa^2/m_Q=-m_Q\alpha_C^2/4\).  The normalization follows from
\[
  4\pi
  \left(\frac{\kappa^3}{\pi}\right)
  \int_0^\infty r^2\ee^{-2\kappa r}\,\dd r
  =
  4\kappa^3\frac{2}{(2\kappa)^3}=1,
\]
and the value at the origin is the square of the normalization constant.
The mass formula is then the definition of the spin-averaged singlet energy
coordinate in this leading Hamiltonian.  This calculation does not define the
physical charmonium or bottomonium spectrum by itself: it computes the first
term in a declared pNRQCD expansion whose mass scheme, matching scale, and
omitted operators must be specified.

\paragraph{Quarkonium fine and hyperfine spin algebra.}
Fix a quarkonium radial level and orbital angular momentum \(L\) in a
pNRQCD specification.  Let
\[
  \mathbf S_Q=\frac12\boldsymbol\sigma_Q,\qquad
  \mathbf S_{\bar Q}=\frac12\boldsymbol\sigma_{\bar Q},\qquad
  \mathbf S=\mathbf S_Q+\mathbf S_{\bar Q},\qquad
  \mathbf J=\mathbf L+\mathbf S .
\]
Assume that, to first order in the spin-dependent part of the matched
Hamiltonian, the restriction to this radial-orbital subspace has the form
\begin{equation}
  H_{nL}^{\rm sd}
  =
  A_{nL}\,\mathbf S_Q\cdot\mathbf S_{\bar Q}
  +
  B_{nL}\,\mathbf L\cdot\mathbf S
  +
  C_{nL}\,\mathcal T_{LS},
  \label{eq:qcd-quarkonium-spin-dependent-hamiltonian}
\end{equation}
where \(A_{nL},B_{nL},C_{nL}\) are radial matrix elements or matching
coefficients in the chosen scheme, and \(\mathcal T_{LS}\) is the rotational
scalar obtained by coupling a traceless rank-two orbital tensor to a
traceless rank-two spin tensor.  Then
\begin{equation}
  \mathbf S_Q\cdot\mathbf S_{\bar Q}
  =
  \begin{cases}
  -\frac34, & S=0,\\[2mm]
  \frac14, & S=1,
  \end{cases}
  \label{eq:qcd-quarkonium-spin-spin-eigenvalues}
\end{equation}
and, on the triplet \(S=1\) states,
\begin{equation}
  \mathbf L\cdot\mathbf S
  =
  \frac12\bigl(J(J+1)-L(L+1)-2\bigr)
  =
  \begin{cases}
  -(L+1), & J=L-1,\\
  -1, & J=L,\\
  L, & J=L+1 .
  \end{cases}
  \label{eq:qcd-quarkonium-spin-orbit-eigenvalues}
\end{equation}
If \(M(^1L_L)\) denotes the singlet mass and \(M(^3L_J)\) the triplet masses
computed to this first order from a common spin-independent radial level
\(E_{nL}^{(0)}\), then the triplet centroid obeys
\begin{equation}
  \overline M(^3L)
  :=
  \frac{\sum_{J=L-1}^{L+1}(2J+1)M(^3L_J)}{3(2L+1)}
  =
  E_{nL}^{(0)}+\frac14 A_{nL},
  \label{eq:qcd-quarkonium-triplet-centroid}
\end{equation}
with the obvious omission of the nonexistent \(J=L-1\) term when \(L=0\).
The centroid hyperfine difference is therefore
\begin{equation}
  \overline M(^3L)-M(^1L_L)=A_{nL}.
  \label{eq:qcd-quarkonium-hyperfine-centroid}
\end{equation}
The two heavy spins are spin-\(\frac12\) representations.  Since
\[
  \mathbf S^2
  =
  \mathbf S_Q^2+\mathbf S_{\bar Q}^2
  +2\mathbf S_Q\cdot\mathbf S_{\bar Q},
  \qquad
  \mathbf S_Q^2=\mathbf S_{\bar Q}^2=\frac34,
\]
one has
\[
  \mathbf S_Q\cdot\mathbf S_{\bar Q}
  =
  \frac12\left(S(S+1)-\frac32\right),
\]
which gives \eqref{eq:qcd-quarkonium-spin-spin-eigenvalues} for
\(S=0,1\).  Similarly
\[
  \mathbf J^2
  =
  \mathbf L^2+\mathbf S^2+2\mathbf L\cdot\mathbf S
\]
gives the first equality in
\eqref{eq:qcd-quarkonium-spin-orbit-eigenvalues}; substituting
\(S=1\) and \(J=L-1,L,L+1\) gives the displayed three values.

It remains to check the centroid.  The trace of \(\mathbf L\cdot\mathbf S\)
on the tensor product of the orbital \(L\) space and the triplet spin space
vanishes:
\[
  \operatorname{Tr}_{L\otimes S=1}(\mathbf L\cdot\mathbf S)
  =
  \sum_i \operatorname{Tr}_{L}(L_i)\operatorname{Tr}_{S=1}(S_i)=0,
\]
because every generator of an irreducible \(SU(2)\) representation is
traceless.  The operator \(\mathcal T_{LS}\) is a contraction of traceless
rank-two tensors, so its trace on the same tensor product also vanishes.  The
weighted sum
\(\sum_J(2J+1)M(^3L_J)\) is precisely this trace after decomposition into
irreducible \(J\) summands.  Therefore only the spin-independent energy and
the triplet spin-spin eigenvalue \(\frac14A_{nL}\) survive in the centroid,
proving \eqref{eq:qcd-quarkonium-triplet-centroid}.  The singlet has
\(\mathbf L\cdot\mathbf S=0\), no rank-two spin tensor, and spin-spin
eigenvalue \(-\frac34\); subtracting its first-order mass from the triplet
centroid gives \eqref{eq:qcd-quarkonium-hyperfine-centroid}.
The spin algebra above is the part of quarkonium spectroscopy that follows
from angular momentum and the matched operator form alone.  The numerical
values of
\(A_{nL},B_{nL},C_{nL}\) are QCD matching data.  In weak-coupling pNRQCD they
are computed from perturbative potentials and wavefunctions in a stated mass
scheme; in strong-coupling pNRQCD they are Wilson-loop and chromofield
insertion matrix elements.  In either case the symbols
``hyperfine splitting'' and ``fine structure'' refer to the spectral
differences defined by
\eqref{eq:qcd-quarkonium-spin-orbit-eigenvalues}--\eqref{eq:qcd-quarkonium-hyperfine-centroid},
rather than to a freestanding constituent-potential model.

\paragraph{Leading weak-coupling \(1S\) hyperfine coordinate.}
In the weak-coupling \(1S\) chart above, retain the leading
chromomagnetic contact operator
\begin{equation}
  H_{\rm ss}^{1S}
  =
  \frac{8\pi\alpha_C\,c_F(\mu)^2}{3m_Q^2}\,
  \mathbf S_Q\cdot\mathbf S_{\bar Q}\,
  \delta^{(3)}(\mathbf r),
  \label{eq:qcd-pnrqcd-contact-spin-spin}
\end{equation}
where \(c_F(\mu)\) is the chromomagnetic matching coefficient in the same
short-distance mass and coupling scheme as \(m_Q\) and \(\alpha_C\).  To first
order in this operator,
\begin{equation}
  M(^3S_1)-M(^1S_0)
  =
  \frac{8\pi\alpha_C\,c_F(\mu)^2}{3m_Q^2}
  |\psi_{1S}(0)|^2
  =
  \frac{m_Q\,c_F(\mu)^2\,\alpha_C^4}{3}.
  \label{eq:qcd-one-s-hyperfine-contact-result}
\end{equation}
The displayed number is a leading coordinate in the declared weak-coupling
pNRQCD scheme.  It is not by itself a precision prediction for charmonium or
bottomonium until the mass scheme, higher-order potentials, ultrasoft terms,
nonperturbative corrections, and continuum-limit uncertainties have been
specified.
First-order perturbation theory in the normalized Coulombic \(1S\) state gives
\[
  \langle H_{\rm ss}^{1S}\rangle_{S}
  =
  \frac{8\pi\alpha_C\,c_F(\mu)^2}{3m_Q^2}
  |\psi_{1S}(0)|^2\,
  \langle\mathbf S_Q\cdot\mathbf S_{\bar Q}\rangle_S .
\]
For the triplet \(S=1\) the spin expectation value is \(1/4\), and for the
singlet \(S=0\) it is \(-3/4\), as shown in
\eqref{eq:qcd-quarkonium-spin-spin-eigenvalues}.  Their difference is one, so
the triplet-singlet mass splitting is exactly the contact coefficient times
\(|\psi_{1S}(0)|^2\).  Substituting
\eqref{eq:qcd-coulombic-one-s-energy-origin},
\[
  |\psi_{1S}(0)|^2=\frac{(m_Q\alpha_C)^3}{8\pi},
\]
gives the second equality in
\eqref{eq:qcd-one-s-hyperfine-contact-result}.  The proof uses only the
declared pNRQCD contact operator and the Coulombic \(1S\) wavefunction; all
other terms belong to the stated remainder of the weak-coupling chart.

\begin{controlledapproximation}[Quarkonium scale separation]
\label{ca:nrqcd-pnrqcd-scale-separation}
Let \(v_{\rm rel}\) be the velocity coordinate assigned to a family of
quarkonium states.  The weak-coupling potential expansion assumes
\[
  m_Q\gg m_Qv_{\rm rel}\gg m_Qv_{\rm rel}^2,
  \qquad
  m_Qv_{\rm rel}\gg \Lambda_{\rm QCD},
\]
so that the soft relative-momentum scale can be integrated out
perturbatively.  Strong-coupling pNRQCD instead treats the potential as a
nonperturbative matching coefficient defined by static Wilson-loop and
chromofield-insertion matrix elements.  In both regimes a claim of accuracy
requires a norm or class of smeared matrix elements and an explicit remainder
estimate in powers of \(v_{\rm rel}\), \(1/m_Q\), and
\(\Lambda_{\rm QCD}/(m_Qv_{\rm rel})\) when that ratio is used.
\end{controlledapproximation}

\section{Heavy-Quark Mass Coordinates and Static Energies}
\label{sec:qcd-heavy-mass-static-energy}

The mass \(m_Q\) appearing in HQET, NRQCD, and pNRQCD is a coordinate in a
regulated Lagrangian and in its matched effective descriptions.  In a confining
four-dimensional QCD theory an isolated colored quark is not an asymptotic
particle, so a quark pole mass is not a gauge-invariant spectral observable.
Perturbation theory in a fixed gauge nevertheless defines formal mass
coordinates, including an on-shell coordinate order by order.  Such
coordinates are useful only after their infrared sensitivity and their
relation to gauge-invariant quantities have been stated.

The gauge-invariant static quark-antiquark energy is extracted from Wilson
loops.  With a specified line-renormalization prescription, define
\begin{equation}
  E_{\rm stat}(r;\mu)
  =
  -\lim_{T\to\infty}\frac1T
  \log \langle W_{\square}(r,T)\rangle_{\rm ren,\mu},
  \label{eq:qcd-static-energy-renormalized}
\end{equation}
where \(W_{\square}(r,T)\) is the rectangular fundamental Wilson loop of
spatial size \(r\) and time extent \(T\).  The subscript records the
ultraviolet and line-renormalization convention.  Additive constants in
\(E_{\rm stat}\) are not absolute observables; energy differences and matched
hadron masses are.

At distances \(r\) in a weak-coupling matching range, the pNRQCD singlet
potential is a coordinate in the decomposition
\begin{equation}
  E_{\rm stat}(r;\mu)
  =
  2m_{\rm sd}(\mu_f)+V_{\rm sd}(r;\mu,\mu_f)+\Delta_{\rm us}(r;\mu,\mu_f),
  \label{eq:qcd-static-energy-sd-decomposition}
\end{equation}
where \(m_{\rm sd}\) is a short-distance heavy-quark mass coordinate,
\(V_{\rm sd}\) is the corresponding singlet potential coordinate, and
\(\Delta_{\rm us}\) denotes lower-energy contributions retained in pNRQCD.
The auxiliary scale \(\mu_f\) separates contributions assigned to the mass
coordinate from contributions assigned to the potential coordinate.

\begin{remark}[Constant mass-coordinate transformations]
\label{rem:qcd-heavy-mass-potential-shift}
Let \(\delta m\) be a finite scheme change independent of \(r\).  The
coordinate transformation
\begin{equation}
  m_{\rm sd}'=m_{\rm sd}+\delta m,
  \qquad
  V_{\rm sd}'(r)=V_{\rm sd}(r)-2\delta m
  \label{eq:qcd-heavy-mass-potential-shift}
\end{equation}
leaves the static energy \eqref{eq:qcd-static-energy-sd-decomposition}
unchanged when \(\Delta_{\rm us}\) is held fixed.  If a pNRQCD Hamiltonian is
shifted by \(V_{\rm sd}\mapsto V_{\rm sd}-2\delta m\), every singlet
eigenvalue \(E_n\) is shifted by \(-2\delta m\), and
\(2m_{\rm sd}+E_n\) is unchanged.
\par
Substitution into \eqref{eq:qcd-static-energy-sd-decomposition} gives
\[
  2(m_{\rm sd}+\delta m)+(V_{\rm sd}-2\delta m)+\Delta_{\rm us}
  =
  2m_{\rm sd}+V_{\rm sd}+\Delta_{\rm us}.
\]
For a Hamiltonian \(H_{\rm sd}=T+V_{\rm sd}\), the transformed Hamiltonian is
\(H_{\rm sd}'=H_{\rm sd}-2\delta m\).  If
\(H_{\rm sd}\psi_n=E_n\psi_n\), then
\(H_{\rm sd}'\psi_n=(E_n-2\delta m)\psi_n\).  Therefore
\[
  2(m_{\rm sd}+\delta m)+(E_n-2\delta m)=2m_{\rm sd}+E_n .
\]
The statement is finite-dimensional spectral algebra applied to the
regularized Hamiltonian, and then to any limit in which the relevant
eigenvalue exists.
\end{remark}

One useful short-distance coordinate is the potential-subtracted mass.  Start
from a formal perturbative pole coordinate \(m_{\rm pole}\) and define
\begin{equation}
  m_{\rm PS}(\mu_f)
  =
  m_{\rm pole}-\delta m_{\rm PS}(\mu_f),
  \qquad
  \delta m_{\rm PS}(\mu_f)
  =
  -\frac12
  \int_{|\mathbf q|<\mu_f}
  \frac{\dd^3\mathbf q}{(2\pi)^3}\,
  \widetilde V(\mathbf q),
  \label{eq:qcd-ps-mass-definition}
\end{equation}
where \(\widetilde V(\mathbf q)\) is the perturbative momentum-space singlet
potential in the same convention.  At tree level
\[
  \widetilde V^{(0)}(\mathbf q)=-\frac{g^2C_F}{|\mathbf q|^2}.
\]
Consequently
\begin{equation}
  \delta m_{\rm PS}^{(0)}(\mu_f)
  =
  \frac{g^2C_F}{4\pi^2}\,\mu_f .
  \label{eq:qcd-ps-leading-subtraction}
\end{equation}
In the common half-trace convention this is the same expression because
\(g_{\rm ht}^2C_F^{\rm ht}=g^2C_F\).

Using spherical coordinates in momentum space,
\[
  \int_{|\mathbf q|<\mu_f}
  \frac{\dd^3\mathbf q}{(2\pi)^3}\,\frac1{|\mathbf q|^2}
  =
  \frac{4\pi}{(2\pi)^3}\int_0^{\mu_f}\dd q
  =
  \frac{\mu_f}{2\pi^2}.
\]
Multiplying by \(g^2C_F/2\), as dictated by
\eqref{eq:qcd-ps-mass-definition} and
\(\widetilde V^{(0)}=-g^2C_F/|\mathbf q|^2\), gives
\eqref{eq:qcd-ps-leading-subtraction}.

\begin{remark}[Renormalon language in the heavy-quark sector]
\label{rem:qcd-heavy-renormalon-status}
The leading infrared sensitivity of the formal perturbative pole coordinate
and the additive constant in the perturbative static potential are tied to the
same long-distance momentum region.  In perturbative large-order language this
is described as a cancellation of the leading static-energy renormalon between
\(2m_{\rm pole}\) and \(V_{\rm pole}(r)\).  The mathematically precise
statement used here is the finite coordinate transformation
\eqref{eq:qcd-heavy-mass-potential-shift} and the definition of
short-distance coordinates such as \eqref{eq:qcd-ps-mass-definition}.  A claim
about a Borel singularity, its residue, or its cancellation is a claim about a
specified perturbative expansion and summation prescription, not a
nonperturbative definition of the quark mass.
\end{remark}

\begin{controlledapproximation}[Mass schemes in quarkonium predictions]
\label{ca:qcd-heavy-mass-scheme-control}
A quarkonium prediction based on pNRQCD must state the mass coordinate, the
potential coordinate, the subtraction scale \(\mu_f\) when one is used, and the
class of matrix elements or energy levels being approximated.  Replacing
\((m,V)\) by a short-distance pair \((m',V')\) is exact at the level of the
matched Hamiltonian if \eqref{eq:qcd-heavy-mass-potential-shift} is imposed.
Truncating perturbative series introduces residual \(\mu_f\)- and
scheme-dependence; the size of that dependence is an error diagnostic only
after the truncation order, velocity counting, and ultrasoft treatment have
been fixed.
\end{controlledapproximation}

\section{Color and Physical External States}

The gauge-fixed quark and gluon fields are useful local variables in
perturbation theory.  Physical states are BRST cohomology classes, and local
physical operators are represented by gauge-invariant or BRST-closed
operators.  This has an immediate consequence for scattering: LSZ reduction
applied to the elementary quark or gluon fields does not by itself define a
physical hadron \(S\)-matrix.  Physical external states in QCD are created by
color-singlet operators, such as
\[
  \bar q_I q_J,\qquad
  \bar q_I\gamma^\mu q_J,\qquad
  \epsilon_{i_1\cdots i_{N_c}}
  q^{i_1}_{I_1}\cdots q^{i_{N_c}}_{I_{N_c}},
  \qquad
  \operatorname{tr}(F_{\mu\nu}F^{\mu\nu}).
\]
The first three are mesonic current, vector-current, and baryonic examples;
the last is a gluonic example.

Confinement, in the spectral sense used here, is the assertion that isolated
finite-energy asymptotic particles carry no color.  In four-dimensional QCD
this assertion is a nonperturbative physical input, supported by the observed
hadron spectrum and by lattice evidence, rather than a theorem presently
derived from the continuum Lagrangian.  The monograph uses this precise
spectral hypothesis when discussing QCD phenomenology: the asymptotic states
are hadrons, while the short-distance operator expansion is expressed in terms
of quark and gluon fields because those are the local variables of the
renormalized gauge theory.

\begin{openproblem}[Asymptotic completeness of physical QCD]
\label{op:qcd-asymptotic-completeness}
Starting from a nonperturbative four-dimensional QCD Hilbert space
\(\mathcal H_{\rm phys}\), its unitary translation representation, and its
gauge-invariant local observable algebra, construct the physical incoming and
outgoing scattering theory.  The required asymptotic space
\[
  \mathcal H_{\rm as}^{\rm QCD}
\]
should be built from stable color-singlet Wigner particles in the relevant
superselection sectors: mesons, baryons, glueballs in pure Yang--Mills
subsectors, and, in chiral limits, the massless Goldstone modes with their
long-distance scattering structure.  Unstable hadronic resonances should enter
through the analytic structure of the color-singlet scattering operator rather
than as ordinary external vectors.  The wave operators
\[
  \Omega_{\rm in/out}^{\rm QCD}:
  \mathcal H_{\rm as}^{\rm QCD}\longrightarrow \mathcal H_{\rm phys}
\]
should be isometric, should intertwine translations, and should have common
closed range equal to the finite-energy physical sector under consideration:
\[
  \overline{\operatorname{Ran}\Omega_{\rm in}^{\rm QCD}}
  =
  \overline{\operatorname{Ran}\Omega_{\rm out}^{\rm QCD}}
  =
  \mathcal H_{\rm phys}^{\rm sector}.
\]
One must also construct the detector limits, such as the energy-flow
operators below, on the same scattering domain and prove that the inclusive
calorimetric observables used in high-energy QCD are obtained from this
Hilbert-space scattering theory.

This problem is logically stronger than asymptotic freedom, factorization, a
mass gap in pure Yang--Mills theory, or the spectral confinement hypothesis
used above.  Asymptotic freedom controls short-distance operator products and
partonic coefficient functions.  Factorization gives a perturbative expansion
for selected inclusive observables after the relevant operator matrix elements
have been defined.  Asymptotic completeness is a large-time statement about
the range of physical wave operators in the gauge-invariant Hilbert space.
\end{openproblem}

\section{Hadron Spectroscopy as Ab Initio QCD}
\label{sec:qcd-hadron-spectroscopy-ab-initio}

\subsection{Spectral Sources and Finite-Volume Data}
\label{subsec:qcd-spectroscopy-sources-finite-volume}

Hadron spectroscopy is not a list of model names assigned to observed peaks.
It is a collection of spectral and scattering statements about the
gauge-invariant Hilbert space of QCD.  A spectroscopic claim must therefore
state which Hilbert-space object is being computed.  Stable hadrons are
isolated eigenvalues of the mass operator \(M^2=P_\mu P^\mu\) in color-singlet
superselection sectors.  Resonances are poles of color-singlet scattering
amplitudes on specified nonphysical sheets, as in
Chapter~\ref{ch:volume-ii-resonances-dressed-propagators}; they are not
additional vectors in the asymptotic Hilbert space.  Finite-volume energy
levels are regulator observables whose infinite-volume interpretation requires
a quantization map, not a direct equality between a box eigenvalue and a
resonance mass.

Let \(I,J\) denote flavor labels and let \(i,j\) be color labels.  Typical
local color-singlet sources are
\begin{align}
  M_{\Gamma,IJ}(x)
  &=
  \bar q_I(x)\Gamma q_J(x),
  \label{eq:qcd-meson-local-source}\\
  B_{\alpha_1\cdots\alpha_{N_c};I_1\cdots I_{N_c}}(x)
  &=
  \epsilon_{i_1\cdots i_{N_c}}\,
  q^{i_1}_{\alpha_1 I_1}(x)\cdots
  q^{i_{N_c}}_{\alpha_{N_c} I_{N_c}}(x),
  \label{eq:qcd-baryon-local-source}\\
  G_{\mathcal P}(x)
  &=
  \operatorname{tr}\mathcal P(F_{\mu\nu},D_\rho)(x).
  \label{eq:qcd-glueball-local-source}
\end{align}
Here \(\Gamma\) is a Dirac matrix, possibly with covariant derivatives
inserted and with Lorentz indices projected to a definite spin channel, and
\(\mathcal P\) is a gauge-covariant polynomial whose color indices are traced
using the convention fixed in \eqref{eq:qcd-casimir-definitions}.  Nonlocal
sources are allowed, but their Wilson lines, smearing kernels, and endpoint
renormalizations are part of the operator definition.  For example, a hybrid
meson source may be represented by
\[
  \bar q_I(x)\Gamma^{\mu\nu} F_{\mu\nu}(x) q_J(x),
\]
or by a point-split bilinear with a Wilson-line path and a field-strength
insertion on that path.  These are different regulator coordinates unless a
continuum matching relation has been supplied.

\begin{definition}[Hadron spectral extraction setup]
\label{def:qcd-hadron-spectral-extraction-setup}
A QCD hadron spectral extraction setup consists of:
\begin{enumerate}
  \item the quark masses, gauge group, global form, spacetime volume, boundary
  conditions, and regulator action;
  \item a list of gauge-invariant interpolating operators, including their
  flavor, parity, charge conjugation, spin or cubic-group channel, smearing,
  and renormalization prescriptions;
  \item the Euclidean correlation matrix
  \[
    C_{AB}(t)=
    \langle \mathcal O_A(t)\mathcal O_B^\dagger(0)\rangle
    -\langle \mathcal O_A\rangle\langle\mathcal O_B^\dagger\rangle ;
  \]
  \item a declared estimator of finite-volume color-singlet energies from the
  transfer matrix or Euclidean correlators, including its source basis, fit
  window, covariance treatment, and stability diagnostics;
  \item a map from finite-volume energies to infinite-volume observables:
  isolated stable masses for one-particle states, and scattering amplitudes or
  pole positions for resonances;
  \item continuum, chiral, heavy-quark, and infinite-volume extrapolations,
  with the systematics stated separately.
\end{enumerate}
The setup is called \emph{ab initio} only when the matrix elements are defined
from the QCD path integral or Hilbert space itself.  A constituent model may
be useful as a coordinate chart or mnemonic, but it is not an ab initio
definition of the spectrum.
\end{definition}

\begin{definition}[Operator-sector spectroscopy chart]
\label{def:qcd-operator-sector-spectroscopy-chart}
Fix a finite-volume regulator and a symmetry channel \(\Gamma\).  Let
\(\mathcal V_\Gamma\) be a finite-dimensional vector space of renormalized
gauge-invariant interpolating operators in that channel, modulo operators
whose matrix elements vanish on the finite-volume Hilbert subspace being
sampled.  A spectroscopy chart is a choice of ordered basis
\((\mathcal O_1,\ldots,\mathcal O_m)\) of \(\mathcal V_\Gamma\), together
with any declared decomposition
\begin{equation}
  \mathcal V_\Gamma
  =
  \mathcal V_{\rm loc}\oplus
  \mathcal V_{\rm two\text{-}hadron}\oplus
  \mathcal V_{\rm flux}\oplus\cdots .
  \label{eq:qcd-spectroscopy-source-decomposition}
\end{equation}
The summands in \eqref{eq:qcd-spectroscopy-source-decomposition} are source
subspaces, not Hilbert-space superselection sectors.  For example,
\(\mathcal V_{\rm loc}\) may contain local \(q\bar q\), baryon, or pure-glue
operators, \(\mathcal V_{\rm two\text{-}hadron}\) may contain products of
separately projected hadron operators, and \(\mathcal V_{\rm flux}\) may
contain point-split bilinears with field-strength insertions or extended
Wilson-line paths.  The decomposition is part of the chart; it is not an
invariant classification of states unless a regulator, renormalization
scheme, and continuum matching prescription have also been supplied.
\end{definition}

For a finite spatial torus and a reflection-positive lattice regulator, the
spectral decomposition has the same form as
\eqref{eq:volxi-glueball-spectral-matrix}:
\[
  C_{AB}(t)=
  \sum_{n\ge0}Z_A^{(n)}\overline{Z_B^{(n)}}\ee^{-E_n t},
  \qquad
  Z_A^{(n)}=\bra{\Omega}\mathcal O_A\ket{n,L}.
\]
The vectors \(\ket{n,L}\) are finite-volume color-singlet eigenvectors.  If
an eigenvalue converges, after removing the center-of-mass motion and taking
the regulator limits, to an isolated one-particle mass below all relevant
thresholds, the resulting number is a stable-hadron mass.  Above threshold the
finite-volume levels instead sample the scattering operator.  In a single
elastic two-particle channel with total momentum zero, particle masses
\(m_1,m_2\), relative momentum \(k\), and energy
\[
  E=\sqrt{m_1^2+k^2}+\sqrt{m_2^2+k^2},
\]
the simplest unmixed partial-wave quantization condition is
\begin{equation}
  \delta_\ell(k)+\phi_{\ell}(q)=n\pi,
  \qquad
  q=\frac{kL}{2\pi},
  \label{eq:qcd-luscher-single-channel-condition}
\end{equation}
where \(\delta_\ell\) is the infinite-volume elastic phase shift and
\(\phi_\ell\) is a known geometric function built from the finite-volume zeta
function.  Equation~\eqref{eq:qcd-luscher-single-channel-condition} is not a
model of a resonance.  It is a finite-volume reconstruction theorem under
elasticity, locality, and regulator hypotheses.  In realistic QCD channels
one must use the coupled-channel determinant version, because cubic symmetry
mixes partial waves and inelastic thresholds are often nearby.

\paragraph{Source coordinates and invariant output.}
The QCD output of the preceding construction is not a preferred source basis.
At finite regulator the invariant spectral data are the color-singlet
transfer-matrix energies and their symmetry labels; after the continuum and
infinite-volume limits they become stable masses, scattering amplitudes, or
sheet-labeled pole data according to the channel.  Source-overlap phrases such
as ``mostly two-hadron'', ``mostly hybrid'', or ``mostly glueball'' are
coordinate statements about the chosen operator family.  They acquire
unambiguous meaning only after the source subspaces in
\eqref{eq:qcd-spectroscopy-source-decomposition}, their renormalization and
smearing prescriptions, and the positive form used to compare overlap
covectors have all been declared.  Changing these source coordinates can
change the quoted overlap fractions without changing the finite-volume energy
or the reconstructed pole.  This chapter therefore records the QCD-specific
operators and the spectral or scattering quantity being extracted, while the
finite-regulator estimator itself is part of the numerical methodology.

\begin{remark}[Spectroscopic names as derived coordinates]
\label{rem:qcd-spectroscopic-names-derived}
In the finite-volume QCD framework above, the invariant output of a stable
sector is an energy level and its quantum numbers, while the invariant output
of a resonance sector is a sheet-labeled pole and residue data after a
finite-volume-to-infinite-volume reconstruction.  Names such as conventional
meson, glueball, hybrid, molecule, tetraquark, or flux-tube excitation are
therefore derived coordinates on source overlaps, pole residues, large-\(N_c\)
scaling, and deformation behavior.  They are not additional definitions of
QCD states.
\end{remark}

\subsection{Elastic Finite-Volume Reconstruction}
\label{subsec:qcd-elastic-finite-volume-reconstruction}

\begin{definition}[Periodic Green function and \(S\)-wave zeta coordinate]
\label{def:qcd-periodic-green-zeta-coordinate}
For a cubic spatial torus of side \(L\), let
\[
  \mathbf p_{\mathbf n}:=\frac{2\pi}{L}\mathbf n,\qquad
  \mathbf n\in\mathbb Z^3,
\]
and let \(k>0\) be such that \(k^2\ne \mathbf p_{\mathbf n}^{\,2}\) for all
\(\mathbf n\).  The periodic Helmholtz Green function is the distribution
\begin{equation}
  G_L(\mathbf r;k)
  :=
  \frac1{L^3}
  \sum_{\mathbf n\in\mathbb Z^3}
  \frac{\ee^{\ii\mathbf p_{\mathbf n}\cdot\mathbf r}}
       {\mathbf p_{\mathbf n}^{\,2}-k^2},
  \qquad
  (\Delta+k^2)G_L(\mathbf r;k)=-\delta_L(\mathbf r),
  \label{eq:qcd-periodic-helmholtz-green}
\end{equation}
where \(\delta_L\) is the periodic delta distribution.  Its singular expansion
at the origin is written
\begin{equation}
  G_L(\mathbf r;k)
  =
  \frac{1}{4\pi|\mathbf r|}
  +
  c_L(k)
  +O(|\mathbf r|),
  \qquad
  \mathbf r\to0.
  \label{eq:qcd-periodic-green-constant-term}
\end{equation}
Equivalently, with \(q=kL/(2\pi)\), define the analytically continued
Luescher zeta coordinate \(\mathcal Z_{00}(1;q^2)\) by
\begin{equation}
  4\pi c_L(k)
  =
  \frac{2}{\sqrt\pi\,L}\,
  \mathcal Z_{00}(1;q^2),
  \qquad
  \mathcal Z_{00}(s;q^2)
  :=
  \frac1{\sqrt{4\pi}}
  \sum_{\mathbf n\in\mathbb Z^3}
  \frac{1}{(\mathbf n^2-q^2)^s}
  \quad(\operatorname{Re}s>\tfrac32),
  \label{eq:qcd-luscher-zeta-coordinate}
\end{equation}
continued meromorphically to \(s=1\).  The singular summands at
\(\mathbf n^2=q^2\) are absent from the elastic interval under discussion;
when a free two-particle level is crossed, the condition is read by the usual
limiting prescription from adjacent \(q^2\).
\end{definition}

\begin{proposition}[\(S\)-wave finite-volume quantization from wave matching]
\label{prop:qcd-swave-luscher-wave-matching}
Consider a single elastic two-particle channel in the center-of-mass frame.
Assume the two stable particles have masses \(m_1,m_2\), the interaction range
in the relative coordinate is contained in \(|\mathbf r|<R\), the box obeys
\(L>2R\), and the finite-volume corrections from virtual particles wrapping
around the torus are bounded by \(C\ee^{-m_{\rm gap}L}\) for some positive
mass scale \(m_{\rm gap}\) in the omitted exchange channel.  In an exact
finite-range model this exponential remainder is absent.  Assume also that
the cubic irrep isolates the \(S\)-wave, so higher partial waves are not
retained in the truncation.  Let \(k\) be the relative momentum determined by
\[
  E=\sqrt{m_1^2+k^2}+\sqrt{m_2^2+k^2}.
\]
Then the finite-volume energy is allowed precisely when
\begin{equation}
  k\cot\delta_0(k)=4\pi c_L(k)
  =
  \frac{2}{\sqrt\pi\,L}\mathcal Z_{00}(1;q^2),
  \qquad q=\frac{kL}{2\pi}.
  \label{eq:qcd-swave-luscher-kcot-form}
\end{equation}
Equivalently,
\begin{equation}
  q\cot\delta_0(k)=\frac1{\pi^{3/2}}\mathcal Z_{00}(1;q^2),
  \qquad
  \delta_0(k)+\phi_0(q)=n\pi,
  \qquad
  \cot\phi_0(q)=
  -\frac{\mathcal Z_{00}(1;q^2)}{\pi^{3/2}q}.
  \label{eq:qcd-swave-luscher-phase-form}
\end{equation}
\end{proposition}

\begin{proof}
Outside the interaction region the relative wave function
\(\psi(\mathbf r)\) satisfies the free Helmholtz equation
\((\Delta+k^2)\psi=0\) and periodic boundary conditions.  In the \(S\)-wave
truncation any such finite-volume solution with a source at the origin is
proportional, in the exterior region \(R<|\mathbf r|<L/2\), to the periodic
Green function \(G_L(\mathbf r;k)\).  The proportionality constant is
irrelevant; the quantization condition is obtained by comparing the ratio of
the regular and singular terms in the short-distance expansion.

The infinite-volume elastic \(S\)-wave outside the interaction range has the
standard real form
\[
  \psi_{\infty}(r)
  =
  A\,\frac{\sin(kr+\delta_0(k))}{kr},
  \qquad r>R .
\]
Expanding this expression at small \(r\) gives
\[
  \psi_{\infty}(r)
  =
  A\left[
    \frac{\sin\delta_0(k)}{kr}
    +
    \cos\delta_0(k)
    +O(r)
  \right].
\]
Thus the ratio of the regular coefficient to the \(1/r\) coefficient is
\begin{equation}
  \frac{A\cos\delta_0(k)}
       {A\sin\delta_0(k)/k}
  =
  k\cot\delta_0(k).
  \label{eq:qcd-swave-boundary-ratio}
\end{equation}
On the other hand, the finite-volume exterior solution proportional to
\eqref{eq:qcd-periodic-helmholtz-green} has expansion
\[
  B\,G_L(\mathbf r;k)
  =
  B\left[
    \frac1{4\pi r}+c_L(k)+O(r)
  \right],
  \qquad r=|\mathbf r|.
  \]
Its regular-to-singular ratio is \(4\pi c_L(k)\).  Matching the exterior
finite-volume solution to the infinite-volume scattering boundary condition
therefore gives the first equality in
\eqref{eq:qcd-swave-luscher-kcot-form}.  The second equality is the zeta
coordinate definition \eqref{eq:qcd-luscher-zeta-coordinate}.

Multiplying the first form by \(L/(2\pi)\) and using \(q=kL/(2\pi)\) gives
the first equation in \eqref{eq:qcd-swave-luscher-phase-form}.  The phase
form follows by defining \(\phi_0\) through
\(\cot\phi_0=-\mathcal Z_{00}(1;q^2)/(\pi^{3/2}q)\):
\[
  \cot\delta_0(k)=-\cot\phi_0(q)
  =
  \cot\bigl(n\pi-\phi_0(q)\bigr),
  \]
which is equivalent to
\(\delta_0(k)+\phi_0(q)=n\pi\) modulo \(\pi\).  The only physics input beyond
the spectral existence of the two-particle level is the replacement of the
interacting QFT outside the range by the elastic scattering boundary
condition, with exponential finite-volume corrections controlled by the
lightest omitted exchange.  Thus the formula is a reconstruction statement
with explicit hypotheses, not a model of the resonance.
\end{proof}

\subsection{Single- and Coupled-Channel Pole Coordinates}
\label{subsec:qcd-single-coupled-channel-pole-coordinates}

For the single-channel finite-volume pole reconstruction target, assume that,
on an open real energy interval \(I\), a QCD channel is
elastic, has two stable color-singlet particles of masses \(m_1,m_2\), and
has been reduced, by a stated cubic-irrep analysis, to a single partial wave
\(\ell\).  Put \(J=\{E^2:E\in I\}\) and
\[
  \rho_{12}(s)=\frac{2k(s)}{\sqrt s},
  \qquad
  E=\sqrt{m_1^2+k(s)^2}+\sqrt{m_2^2+k(s)^2},
\]
for \(s=E^2\in J\), and use the partial-wave normalization
\[
  S_\ell(s)=1+2\ii \rho_{12}(s)t_\ell^{(I)}(s)
  =\ee^{2\ii\delta_\ell(s)} .
\]
Then the physical-sheet elastic amplitude is equivalently written
\begin{equation}
  t_\ell^{(I)}(s)=
  \frac{1}{K_\ell^{-1}(s)-\ii\rho_{12}(s)},
  \qquad
  K_\ell^{-1}(s)=\rho_{12}(s)\cot\delta_\ell(s),
  \label{eq:qcd-single-channel-kmatrix-definition}
\end{equation}
where \(K_\ell^{-1}(s)\) is real on \(J\).  Each finite-volume energy
\(E_n\in I\) satisfying
\eqref{eq:qcd-luscher-single-channel-condition} determines the real-axis
coordinate
\begin{equation}
  K_\ell^{-1}(E_n^2)
  =
  -\rho_{12}(E_n^2)\cot\phi_\ell(q_n),
  \qquad
  q_n=\frac{k(E_n)L}{2\pi}.
  \label{eq:qcd-luscher-kmatrix-sample}
\end{equation}
If \(K_\ell^{-1}\) and \(\rho_{12}\) have the chosen analytic continuations
to a simply connected neighborhood of a candidate resonance pole and no
additional channel threshold crosses that neighborhood, the corresponding
second-sheet pole \(s_\ast\) is a zero of
\begin{equation}
  K_\ell^{-1}(s_\ast)+\ii\rho_{12}(s_\ast)=0 .
  \label{eq:qcd-single-channel-second-sheet-pole}
\end{equation}

This is the algebraic target of a one-channel finite-volume resonance
extraction, not by itself a proof that the required analytic continuation has
been controlled.  Elastic unitarity in the normalization above says
\(S_\ell^\ast S_\ell=1\) on \(I\).  Hence
\[
  t_\ell^{(I)}(s)
  =
  \frac{\ee^{2\ii\delta_\ell(s)}-1}{2\ii\rho_{12}(s)}
  =
  \frac{\ee^{\ii\delta_\ell(s)}\sin\delta_\ell(s)}
       {\rho_{12}(s)}
\]
and therefore
\[
  \bigl(t_\ell^{(I)}(s)\bigr)^{-1}
  =
  \rho_{12}(s)\bigl(\cot\delta_\ell(s)-\ii\bigr).
\]
This proves \eqref{eq:qcd-single-channel-kmatrix-definition} and identifies
the real coordinate \(K_\ell^{-1}=\rho_{12}\cot\delta_\ell\).  At a
finite-volume level satisfying
\(\delta_\ell(k_n)+\phi_\ell(q_n)=n\pi\),
\[
  \cot\delta_\ell(k_n)
  =
  \cot\bigl(n\pi-\phi_\ell(q_n)\bigr)
  =
  -\cot\phi_\ell(q_n),
\]
which gives \eqref{eq:qcd-luscher-kmatrix-sample}.  Across the elastic
unitarity cut the chosen branch of the two-body momentum changes sign, so
\(\rho_{12}\) changes sign while \(K_\ell^{-1}\), being the real analytic
coordinate fixed on the cut, continues through the same germ.  Thus the
amplitude on the adjacent sheet is
\[
  t_\ell^{(II)}(s)=
  \frac{1}{K_\ell^{-1}(s)+\ii\rho_{12}(s)} ,
\]
and its isolated poles are precisely the zeros of the denominator, giving
\eqref{eq:qcd-single-channel-second-sheet-pole}.

A concrete determination of \(s_\ast\) also contains an analytic representation
or parametrization of \(K_\ell^{-1}\), regulator and volume limits for the
input energies, and separate bounds for neglected partial waves and inelastic
channels.

\begin{lemma}[Coupled-channel \(K\)-matrix determinant and sheet continuation]
\label{lem:qcd-coupled-channel-kmatrix-sheet}
Let \(a,b=1,\ldots,N\) label finitely many stable two-particle
color-singlet channels in a fixed set of conserved quantum numbers, and let
\(\alpha,\beta\) label the partial waves and cubic-irrep components retained
in a finite-volume analysis.  On a real interval below the next omitted
threshold, let \(\rho(s)\) be the positive diagonal matrix with entries
\(\rho_a(s)=2k_a(s)/\sqrt s\), repeated on the retained angular-momentum
indices.  With the normalization
\[
  S=1+2\ii\,\rho^{1/2}t^{(I)}\rho^{1/2},
\]
elastic coupled-channel unitarity is equivalent, on the interval, to
\begin{equation}
  t^{(I)}(s)=\bigl(K^{-1}(s)-\ii\rho(s)\bigr)^{-1},
  \qquad
  K^{-1}(s)^\dagger=K^{-1}(s).
  \label{eq:qcd-coupled-channel-kmatrix}
\end{equation}
Suppose, in addition, that the coupled-channel finite-volume reconstruction
theorem has been established for this channel and written in the same
normalization, so that the finite-volume energies in a cubic irrep
\(\Gamma\) obey
\begin{equation}
  \det\!\bigl(K^{-1}(E^2)+F^\Gamma(E,L)\bigr)=0,
  \label{eq:qcd-coupled-channel-luscher-determinant}
\end{equation}
where \(F^\Gamma(E,L)\) is the real geometric matrix determined by the box,
total momentum, masses, and irrep projection.  In the one-channel unmixed
limit \(F^\Gamma=\rho_{12}\cot\phi_\ell\), so
\eqref{eq:qcd-coupled-channel-luscher-determinant} reduces to
\eqref{eq:qcd-luscher-kmatrix-sample}.  If \(\Sigma\) is the diagonal matrix
whose entry is \(+1\) on a channel kept on the physical branch and \(-1\) on
a channel analytically continued through its unitary cut, the corresponding
sheet amplitude is
\begin{equation}
  t^{(\Sigma)}(s)=
  \bigl(K^{-1}(s)-\ii\,\Sigma\rho(s)\bigr)^{-1},
  \qquad
  \det\!\bigl(K^{-1}(s_\ast)-\ii\,\Sigma\rho(s_\ast)\bigr)=0 .
  \label{eq:qcd-coupled-channel-sheet-pole}
\end{equation}
\end{lemma}

\begin{proof}
From \(S^\dagger S=1\) and
\(S=1+2\ii\rho^{1/2}t^{(I)}\rho^{1/2}\), one obtains
\[
  t^{(I)}-t^{(I)\dagger}
  =
  2\ii\,t^{(I)\dagger}\rho\,t^{(I)}.
\]
Multiplying by \(\bigl(t^{(I)\dagger}\bigr)^{-1}\) on the left and by
\(\bigl(t^{(I)}\bigr)^{-1}\) on the right gives
\[
  \bigl(t^{(I)\dagger}\bigr)^{-1}
  -
  \bigl(t^{(I)}\bigr)^{-1}
  =
  2\ii\rho,
  \qquad
  \bigl(t^{(I)}\bigr)^{-1}
  -
  \bigl(t^{(I)\dagger}\bigr)^{-1}
  =
  -2\ii\rho .
\]
Thus \(K^{-1}:=\bigl(t^{(I)}\bigr)^{-1}+\ii\rho\) is Hermitian on the
physical interval, and \eqref{eq:qcd-coupled-channel-kmatrix} follows.  The
finite-volume determinant relation is the stated reconstruction input
expressed in this \(K\)-matrix coordinate.  In the one-channel unmixed
reduction, its zero condition is
\[
  K_\ell^{-1}(E^2)+\rho_{12}(E^2)\cot\phi_\ell(q)=0,
  \qquad
  K_\ell^{-1}(E^2)
  =
  -\rho_{12}(E^2)\cot\phi_\ell(q),
\]
which is exactly \eqref{eq:qcd-luscher-kmatrix-sample}.  Finally, analytic
continuation through a chosen channel cut changes the sign of that channel's
two-body momentum, hence of the corresponding diagonal entry of \(\rho\),
while the \(K^{-1}\) germ is held fixed.  Repeating this sign choice for all
channels gives the matrix \(\Sigma\), the sheet amplitude in
\eqref{eq:qcd-coupled-channel-sheet-pole}, and the determinant pole
condition.
\end{proof}

\begin{definition}[QCD resonance pole data]
\label{def:qcd-resonance-pole-data}
The resonance pole data in a hadronic channel consist of a channel space of stable
color-singlet asymptotic states, a physical-region scattering matrix
\(S_{ab}(s)\), a specified analytic continuation across specified unitary
cuts, and an isolated pole \(s_\ast\) of the continued matrix or of an
equivalent partial-wave amplitude.  The mass and width coordinates
\[
  \sqrt{s_\ast}=M_\ast-\frac{\ii}{2}\Gamma_\ast
\]
are pole coordinates after a branch of the square root has been declared.
For a broad or overlapping resonance, line-shape parameters on the real axis
are not invariant substitutes for \(s_\ast\).
\end{definition}

This definition is the correct language for the \(\rho\), the light scalar
isoscalar pole usually denoted \(f_0(500)\), many excited baryons, and
near-threshold exotic candidates.  The notation ``tetraquark'', ``molecule'',
``hybrid'', or ``glueball component'' is then a statement about operator
overlaps, pole motion under controlled deformations, large-\(N_c\) scaling,
or residue factorization.  It is not a separate definition of the state.

\begin{definition}[Controlled pole-motion chart]
\label{def:qcd-controlled-pole-motion-chart}
Let \(\eta\) be a real coordinate on a family of QCD-like theories or
regulators whose action, operator definitions, stable-channel spectrum, and
renormalization prescription have been specified for \(\eta\) in an open
interval \(I\).  Examples include a quark-mass coordinate, \(1/N_c\) in a
large-\(N_c\) family, a finite-volume boundary condition used as an auxiliary
probe before the infinite-volume limit, or a coupling to a source that is set
to zero after differentiating.  A pole-motion chart in a fixed set of
conserved quantum numbers consists of:
\begin{enumerate}
  \item a channel space \(\mathcal C_\eta\) of stable color-singlet
  asymptotic states for each \(\eta\), with an identification of these
  channel spaces over the interval where no threshold crossing changes the
  channel list;
  \item an analytic family of continued partial-wave amplitudes
  \(t^{(\Sigma)}_{ab}(s;\eta)\) on a declared sheet \(\Sigma\);
  \item an isolated simple pole \(s_\ast(\eta)\) and residue vectors
  \(g_a(\eta)\) satisfying
  \[
    t_{ab}^{(\Sigma)}(s;\eta)
    =
    \frac{g_a(\eta)g_b(\eta)}{s_\ast(\eta)-s}
    +h_{ab}(s;\eta),
  \]
  with \(h_{ab}\) holomorphic near \(s_\ast(\eta)\).
\end{enumerate}
The derivative \(\dd s_\ast/\dd\eta\), the projective residue vector
\([g_a(\eta)]\), and their behavior near thresholds are coordinates on the
declared chart.  They are not intrinsic labels unless the family, channel
identification, and sheet have been included in the statement.
\end{definition}

\subsection{Near-Threshold Effective-Range Coordinates}
\label{subsec:qcd-near-threshold-effective-range-coordinates}

For a shallow \(S\)-wave bound-state effective-range chart, consider one
stable two-hadron channel with reduced mass \(\mu\), threshold
energy set to \(E=0\), and center-of-mass momentum
\(k=\sqrt{2\mu E}\), where the physical branch has \(k>0\) for \(E>0\) and
\(\operatorname{Im}k>0\) below threshold.  Suppose that in an energy disk
whose radius is large compared with the binding energy but small compared
with the first omitted left-hand or inelastic singularity, the elastic
\(S\)-wave amplitude is represented by the effective-range form
\begin{equation}
  f(k)
  =
  \frac{1}{-a^{-1}+\frac12 r_e k^2-\ii k},
  \label{eq:qcd-effective-range-amplitude}
\end{equation}
where \(a\) is the scattering length and \(r_e\) is the effective range in
this normalization.  If \(f\) has a bound-state pole at
\(E=-B=-\kappa^2/(2\mu)\), with \(\kappa>0\), then
\begin{equation}
  \frac1a=\kappa-\frac12 r_e\kappa^2.
  \label{eq:qcd-effective-range-bound-pole-condition}
\end{equation}
Moreover, near \(E=-B\),
\begin{equation}
  f(E)
  =
  -\,\frac{g_{\rm eff}^2}{E+B}+O(1),
  \qquad
  g_{\rm eff}^2
  =
  \frac{\kappa}{\mu(1-r_e\kappa)}.
  \label{eq:qcd-effective-range-bound-residue}
\end{equation}
When \(1-r_e\kappa\ne0\), the dimensionless coordinate
\begin{equation}
  X_{\rm ERE}
  :=
  \frac{a\kappa}{2-a\kappa}
  =
  \frac{1}{1-r_e\kappa}
  \label{eq:qcd-effective-range-compositeness-coordinate}
\end{equation}
is determined by the low-energy pole and the first two effective-range
parameters.  It is a compositeness coordinate in this one-channel
effective-range chart, not a regulator-independent probability attached to a
QCD eigenstate without the hypotheses above.

The bound-state pole lies on the physical sheet at \(k=\ii\kappa\).  The
denominator of \eqref{eq:qcd-effective-range-amplitude} therefore vanishes:
\[
  -a^{-1}
  +\frac12 r_e(\ii\kappa)^2
  -\ii(\ii\kappa)
  =
  -a^{-1}-\frac12 r_e\kappa^2+\kappa=0,
\]
which gives \eqref{eq:qcd-effective-range-bound-pole-condition}.

Let
\[
  D(k)=-a^{-1}+\frac12 r_e k^2-\ii k.
\]
Since \(E=k^2/(2\mu)\), one has
\[
  \frac{\dd D}{\dd E}
  =
  (r_e k-\ii)\frac{\mu}{k}.
\]
Evaluating at \(k=\ii\kappa\) gives
\[
  \left.\frac{\dd D}{\dd E}\right|_{E=-B}
  =
  \frac{\mu(r_e\kappa-1)}{\kappa}.
\]
Thus
\[
  D(E)=\frac{\mu(r_e\kappa-1)}{\kappa}(E+B)+O((E+B)^2),
\]
and
\[
  f(E)=
  \frac{\kappa}{\mu(r_e\kappa-1)}\frac1{E+B}+O(1)
  =
  -\,\frac{\kappa}{\mu(1-r_e\kappa)}\frac1{E+B}+O(1),
\]
which is \eqref{eq:qcd-effective-range-bound-residue}.  Finally,
\eqref{eq:qcd-effective-range-bound-pole-condition} gives
\[
  a\kappa=\frac{1}{1-\frac12 r_e\kappa}.
\]
Substituting this expression into \(a\kappa/(2-a\kappa)\) yields
\[
  \frac{a\kappa}{2-a\kappa}
  =
  \frac{1}{1-r_e\kappa},
\]
proving \eqref{eq:qcd-effective-range-compositeness-coordinate}.

The calculation is deliberately local.  It does not classify every
near-threshold QCD pole.  In coupled channels, or when the omitted range scale
is not separated from \(\kappa^{-1}\), the residue vector and pole motion of
Definition~\ref{def:qcd-controlled-pole-motion-chart} are the appropriate
coordinates.  A statement that a pole is molecular, compact, hybrid, or
predominantly glueball-like must then be translated into source overlaps,
residues, large-\(N_c\) behavior, or a specified deformation path.

The light pseudoscalar sector has additional structure because chiral symmetry
and the anomaly constrain the low-energy coordinates.  Chapter~\ref{ch:volume-ii-21}
derives the pion effective theory, the Gell-Mann--Oakes--Renner relation, and
the Witten--Veneziano large-\(N_c\) formula for the singlet mass.  The vector
and scalar resonances require the scattering-amplitude language above.

\subsection{Light-Meson Scattering and Roy Coordinates}
\label{subsec:qcd-light-meson-roy-coordinates}

\begin{definition}[\(\pi\pi\) isospin amplitudes]
\label{def:qcd-pipi-isospin-amplitudes}
For equal pion masses let \(s+t+u=4m_\pi^2\) and write the
\(SU(2)_V\)-invariant four-pion amplitude as in
\eqref{eq:pion-scattering-isospin-decomposition},
\[
  \mathcal A_{ab\to cd}
  =
  \delta_{ab}\delta_{cd}A(s,t,u)
  +\delta_{ac}\delta_{bd}A(t,s,u)
  +\delta_{ad}\delta_{bc}A(u,t,s).
\]
The scalar function \(A\) is symmetric in its last two arguments; this is the
Bose symmetry obtained by interchanging the two pions in an equal-mass
initial or final pair while keeping the tensor decomposition above fixed.
The \(s\)-channel isospin amplitudes are
\begin{equation}
  \begin{pmatrix}T^0\\ T^1\\ T^2\end{pmatrix}
  =
  \begin{pmatrix}
    3&1&1\\
    0&1&-1\\
    0&1&1
  \end{pmatrix}
  \begin{pmatrix}
    A(s,t,u)\\ A(t,s,u)\\ A(u,t,s)
  \end{pmatrix}.
  \label{eq:qcd-pipi-isospin-amplitude-matrix}
\end{equation}
Equivalently,
\[
  T^0=3A(s,t,u)+A(t,s,u)+A(u,t,s),\quad
  T^1=A(t,s,u)-A(u,t,s),\quad
  T^2=A(t,s,u)+A(u,t,s).
\]
\end{definition}

\paragraph{Crossing matrices in the \(\pi\pi\) isospin basis.}
With \(\mathbf T:=(T^0,T^1,T^2)^T\), crossing \(s\leftrightarrow t\) and
\(s\leftrightarrow u\) is
\begin{equation}
  \mathbf T(t,s,u)=C_{st}\mathbf T(s,t,u),
  \qquad
  \mathbf T(u,t,s)=C_{su}\mathbf T(s,t,u),
  \label{eq:qcd-pipi-crossing-matrices}
\end{equation}
where
\begin{equation}
  C_{st}=
  \begin{pmatrix}
    \frac13&1&\frac53\\
    \frac13&\frac12&-\frac56\\
    \frac13&-\frac12&\frac16
  \end{pmatrix},
  \qquad
  C_{su}=
  \begin{pmatrix}
    \frac13&-1&\frac53\\
    -\frac13&\frac12&\frac56\\
    \frac13&\frac12&\frac16
  \end{pmatrix}.
  \label{eq:qcd-pipi-crossing-matrix-values}
\end{equation}
The matrices in \eqref{eq:qcd-pipi-crossing-matrix-values} are part of the
normalization of the Roy coordinates used below.
Let
\[
  \mathbf A:=
  \begin{pmatrix}
    A(s,t,u)\\ A(t,s,u)\\ A(u,t,s)
  \end{pmatrix},
  \qquad
  M:=
  \begin{pmatrix}
    3&1&1\\
    0&1&-1\\
    0&1&1
  \end{pmatrix}.
\]
Equation~\eqref{eq:qcd-pipi-isospin-amplitude-matrix} is
\(\mathbf T=M\mathbf A\), and
\[
  M^{-1}=
  \begin{pmatrix}
    \frac13&0&-\frac13\\
    0&\frac12&\frac12\\
    0&-\frac12&\frac12
  \end{pmatrix}.
\]
The exchange \(s\leftrightarrow t\) sends
\(\mathbf A\) to \(P_{st}\mathbf A\), where \(P_{st}\) interchanges the first
two components, and \(s\leftrightarrow u\) sends \(\mathbf A\) to
\(P_{su}\mathbf A\), where \(P_{su}\) interchanges the first and third
components.  Therefore \(C_{st}=MP_{st}M^{-1}\) and
\(C_{su}=MP_{su}M^{-1}\), which gives
\eqref{eq:qcd-pipi-crossing-matrix-values}.  In particular
\(C_{st}^2=C_{su}^2=1\), as required by applying the same crossing operation
twice.

For \(\pi\pi\) scattering, write the isospin partial waves as \(t_\ell^I(s)\)
by
\begin{equation}
  T^I(s,t(s,z),u(s,z))
  =
  32\pi\sum_{\ell=0}^{\infty}(2\ell+1)P_\ell(z)\,t_\ell^I(s),
  \qquad
  t(s,z)=-\frac{s-4m_\pi^2}{2}(1-z),
  \label{eq:qcd-pipi-partial-wave-normalization}
\end{equation}
with projection
\[
  t_\ell^I(s)=
  \frac1{64\pi}\int_{-1}^1
  P_\ell(z)T^I(s,t(s,z),u(s,z))\,\dd z .
\]
In this normalization physical-region elastic unitarity is
\[
  \operatorname{Im}t_\ell^I(s)
  =
  \rho_{\pi\pi}(s)|t_\ell^I(s)|^2,
  \qquad
  \rho_{\pi\pi}(s)=\sqrt{1-\frac{4m_\pi^2}{s}},
\]
below the first inelastic threshold.  The \(S\)-wave scattering-length
coordinates used in this section are
\[
  a_0^0:=t_0^0(4m_\pi^2),\qquad
  a_0^2:=t_0^2(4m_\pi^2).
\]

\begin{definition}[Roy input for \(\pi\pi\) scattering]
\label{def:qcd-pipi-roy-input}
Roy input on a real interval \(J\subset[4m_\pi^2,s_{\rm max})\) consists of:
\begin{enumerate}
  \item the pion mass, the isospin amplitudes \(\mathbf T\), and the partial
  wave normalization \eqref{eq:qcd-pipi-partial-wave-normalization};
  \item a domain of fixed-\(t\) analyticity in which twice-subtracted
  Cauchy representations are valid and the boundary values on the right- and
  left-hand cuts exist as tempered distributions;
  \item the crossing matrices \eqref{eq:qcd-pipi-crossing-matrix-values};
  \item subtraction constants \(a_0^0,a_0^2\);
  \item high-energy and high-partial-wave absorptive input sufficient for the
  two subtractions and for the convergence of the partial-wave projections.
\end{enumerate}
The subtraction polynomial projected to partial waves is
\begin{align}
  k_0^0(s;a_0^0,a_0^2)
  &=
  a_0^0
  +
  \frac{s-4m_\pi^2}{12m_\pi^2}\,(2a_0^0-5a_0^2),
  \label{eq:qcd-roy-subtraction-polynomial-00}\\
  k_1^1(s;a_0^0,a_0^2)
  &=
  \frac{s-4m_\pi^2}{72m_\pi^2}\,(2a_0^0-5a_0^2),
  \label{eq:qcd-roy-subtraction-polynomial-11}\\
  k_0^2(s;a_0^0,a_0^2)
  &=
  a_0^2
  -
  \frac{s-4m_\pi^2}{24m_\pi^2}\,(2a_0^0-5a_0^2),
  \label{eq:qcd-roy-subtraction-polynomial-02}
\end{align}
and \(k_\ell^I=0\) for the other partial waves in the twice-subtracted
polynomial.  This normalization fixes what is meant below by the Roy kernels
\(K_{\ell\ell'}^{II'}\).
\end{definition}

\begin{proposition}[Roy projection from fixed-\(t\) analyticity and crossing]
\label{prop:qcd-roy-projection-construction}
Assume the Roy input of Definition~\ref{def:qcd-pipi-roy-input}.  Then, for
\(s\in J\), the real part of each retained partial wave obeys
\begin{equation}
  \operatorname{Re}t_\ell^I(s)
  =
  k_\ell^I(s;a_0^0,a_0^2)
  +
  \sum_{I',\ell'}\int_{4m_\pi^2}^{\infty}\dd s'\,
  K_{\ell\ell'}^{II'}(s,s')\,\operatorname{Im}t_{\ell'}^{I'}(s').
  \label{eq:qcd-roy-equation-projected}
\end{equation}
More explicitly, if the twice-subtracted fixed-\(t\) representation is written
after crossing the left-hand cut into the form
\begin{equation}
  \operatorname{Re}T^I(s,t(s,z))
  =
  P^I(s,z;a_0^0,a_0^2)
  +
  32\pi
  \sum_{I',\ell'}(2\ell'+1)
  \int_{4m_\pi^2}^{\infty}\dd s'\,
  R_{\ell'}^{II'}(s,z;s')\,
  \operatorname{Im}t_{\ell'}^{I'}(s'),
  \label{eq:qcd-fixed-t-roy-resolvent-form}
\end{equation}
then
\begin{equation}
  k_\ell^I(s)
  =
  \frac1{64\pi}\int_{-1}^1
  P_\ell(z)P^I(s,z)\,\dd z,
  \qquad
  K_{\ell\ell'}^{II'}(s,s')
  =
  \frac{2\ell'+1}{2}
  \int_{-1}^1
  P_\ell(z)R_{\ell'}^{II'}(s,z;s')\,\dd z .
  \label{eq:qcd-roy-kernel-projection-definition}
\end{equation}
The functions \(R_{\ell'}^{II'}\), hence the kernels
\(K_{\ell\ell'}^{II'}\), are determined by the Cauchy denominators, the two
subtractions, the crossing matrices \(C_{st},C_{su}\), and the Legendre
projection of the absorptive parts.
\end{proposition}

\begin{proof}
Fix \(t\) in the declared fixed-\(t\) analyticity domain.  The twice-subtracted
Cauchy representation expresses the boundary value of \(\mathbf T(s,t)\) as a
subtraction polynomial plus integrals over the discontinuities on the
right-hand and left-hand cuts.  By crossing, the left-hand discontinuity is
rewritten as a right-hand discontinuity of the crossed isospin vector,
multiplied by \(C_{st}\) or \(C_{su}\) according to which crossed channel has
become physical.  This is the step where
\eqref{eq:qcd-pipi-crossing-matrices} enters; without it the left-hand cut is
not a QCD input in the same isospin coordinates.

For \(s'\ge4m_\pi^2\), expand the absorptive part in partial waves,
\[
  \operatorname{Im}T^{I'}(s',t(s',z'))
  =
  32\pi
  \sum_{\ell'=0}^\infty
  (2\ell'+1)P_{\ell'}(z')\operatorname{Im}t_{\ell'}^{I'}(s').
  \]
Performing the \(z'\)-projection contained in the crossed-channel Cauchy
terms gives the functions \(R_{\ell'}^{II'}(s,z;s')\) in
\eqref{eq:qcd-fixed-t-roy-resolvent-form}.  The convergence assumptions in
the Roy input justify interchanging this partial-wave expansion with the
subtracted dispersion integrals on \(J\).

Finally project the resulting fixed-\(t\) formula in the \(s\)-channel:
\[
  \operatorname{Re}t_\ell^I(s)
  =
  \frac1{64\pi}\int_{-1}^{1}P_\ell(z)
  \operatorname{Re}T^I(s,t(s,z))\,\dd z .
  \]
Substituting \eqref{eq:qcd-fixed-t-roy-resolvent-form} and using the
orthogonality normalization already built into
\eqref{eq:qcd-pipi-partial-wave-normalization} gives exactly
\eqref{eq:qcd-roy-equation-projected} with the projected polynomial and
kernel in \eqref{eq:qcd-roy-kernel-projection-definition}.  The explicit
polynomials \eqref{eq:qcd-roy-subtraction-polynomial-00}--\eqref{eq:qcd-roy-subtraction-polynomial-02}
are the projection of the two independent isospin-even subtraction constants;
the \(I=1\), \(\ell=1\) subtraction is proportional to the same combination
\(2a_0^0-5a_0^2\) because crossing fixes the first \(s-4m_\pi^2\) derivative
of the odd isospin amplitude at threshold.
\end{proof}

\begin{remark}[Status of a Roy pole extraction]
A pole extraction from \eqref{eq:qcd-roy-equation-projected} is controlled
only after the high-energy absorptive input, inelastic channels, matching
point, subtraction constants, and error correlations have been specified.
For the \(\rho\), the relevant pole is obtained by continuing the \(I=1\),
\(\ell=1\) partial wave through the elastic cut.  For the light scalar
isoscalar pole \(f_0(500)\), the corresponding continuation is in the
\(I=0\), \(\ell=0\) wave and is especially sensitive to the subtraction
constants and low-energy crossing constraints.  The same logic governs
Roy--Steiner equations for \(\pi K\), \(\pi N\), and related channels, with
unequal masses and additional crossed-channel partial waves changing the
kernel construction but not the Cauchy-crossing-projection logic.
\end{remark}

\subsection{Light-Meson Pole Charts}
\label{subsec:qcd-light-meson-pole-charts}

\begin{definition}[Concrete light-meson pole chart]
\label{def:qcd-concrete-light-meson-pole-chart}
A concrete light-meson pole statement in this chapter consists of a
color-singlet channel, a partial wave, a reconstruction method, and a
sheet-labeled pole or stable mass.  The minimal chart used for the examples
below is:
\begin{enumerate}[label=(\roman*)]
  \item \(\pi,K,\eta,\eta'\).  The QCD channel is the pseudoscalar flavor
  sector.  The amplitude coordinate is a stable mass, together with mixing
  coordinates where appropriate.  The controlled route uses chiral effective
  theory matched to QCD, lattice masses, and the anomaly constraints of
  Chapter~\ref{ch:volume-ii-20}.
  \item \(\rho\).  The QCD channel is \(I=1\), \(\ell=1\) \(\pi\pi\).  The
  amplitude coordinate is the second-sheet pole of \(t_1^1(s)\).  The
  controlled route is Roy input or finite-volume \(P\)-wave \(\pi\pi\)
  quantization followed by analytic continuation.
  \item \(f_0(500)\).  The QCD channel is \(I=0\), \(\ell=0\) \(\pi\pi\).
  The amplitude coordinate is a broad second-sheet pole of \(t_0^0(s)\).  The
  controlled route is Roy input with subtraction constants and correlated
  high-energy input.
  \item \(K^\ast\).  The QCD channel is \(I=\frac12\), \(P\)-wave \(\pi K\).
  The amplitude coordinate is the second-sheet pole of the unequal-mass
  partial wave.  The controlled route is Roy--Steiner input or finite-volume
  \(\pi K\) quantization.
  \item \(K_0^\ast\) channel.  The QCD channel is \(I=\frac12\), \(S\)-wave
  \(\pi K\).  The amplitude coordinate is nearby pole data, often broad.  The
  controlled route is Roy--Steiner input with explicit treatment of inelastic
  and left-hand input.
  \item Hybrid or exotic \(1^{-+}\) channels.  The QCD channel is specified
  by a finite operator-source family containing \(q\bar qF\) or Wilson-line
  operators.  The amplitude coordinate is pole data plus source-overlap and
  pole-motion coordinates.
  The controlled route supplies color-singlet finite-volume energies from the
  declared source family and then performs finite-volume scattering
  reconstruction.
\end{enumerate}
The entries are not model assignments.  They are instructions for which QCD
operator sector, partial wave, analytic sheet, and reconstruction theorem must
be supplied before the label denotes a spectroscopic output.
\end{definition}

\paragraph{Linear elastic pole coordinate near a narrow vector resonance.}
Consider a one-channel elastic partial wave in the normalization
\[
  t^{(I)}(s)=\frac{1}{K^{-1}(s)-\ii\rho(s)},
  \qquad
  t^{(II)}(s)=\frac{1}{K^{-1}(s)+\ii\rho(s)}.
\]
Let \(M_0^2\) be a real point in the elastic interval such that
\[
  K^{-1}(M_0^2)=0,\qquad
  A:=\left.\frac{\dd K^{-1}}{\dd s}\right|_{M_0^2}>0,
  \qquad
  \rho_0:=\rho(M_0^2)>0,
\]
and suppose that, in a neighborhood whose size is small compared with the
nearest omitted singularity,
\[
  K^{-1}(s)=A(s-M_0^2)+O((s-M_0^2)^2),
  \qquad
  \rho(s)=\rho_0+O(s-M_0^2).
\]
Then the adjacent-sheet pole in this linearized chart is
\begin{equation}
  s_\ast=M_0^2-\ii\,\frac{\rho_0}{A}
  +O\!\left(\frac{\rho_0^2}{A^2}\,
  \sup |(K^{-1})''|+\frac{\rho_0}{A}\sup|\rho'|\right),
  \label{eq:qcd-linear-elastic-pole-coordinate}
\end{equation}
where the suprema are taken on the same neighborhood.  If the pole is
parametrized by
\[
  \sqrt{s_\ast}=M_\ast-\frac{\ii}{2}\Gamma_\ast
\]
and the pole is narrow enough that
\(|s_\ast-M_0^2|\ll M_0^2\), then
\begin{equation}
  M_\ast=M_0+O(\Gamma_\ast^2/M_0),
  \qquad
  \Gamma_\ast=\frac{\rho_0}{A M_0}
  +O(\rho_0^2/A^2M_0^3).
  \label{eq:qcd-linear-elastic-width-coordinate}
\end{equation}
In the exactly linear and constant approximation the second-sheet pole
condition is
\[
  K^{-1}(s_\ast)+\ii\rho(s_\ast)=0
  \quad\Longrightarrow\quad
  A(s_\ast-M_0^2)+\ii\rho_0=0,
\]
which gives \(s_\ast=M_0^2-\ii\rho_0/A\).  The displayed remainder follows
by substituting
\[
  s_\ast=M_0^2-\ii\rho_0/A+\Delta
\]
into the full denominator and solving once for \(\Delta\) by Taylor's theorem
on the neighborhood where the derivative of
\(K^{-1}+\ii\rho\) remains bounded away from zero.  This is an ordinary
one-complex-variable implicit-function estimate; its small parameters are
the quadratic variation of \(K^{-1}\) and the variation of the phase-space
factor across the pole displacement.

Finally,
\[
  (M_\ast-\ii\Gamma_\ast/2)^2
  =
  M_\ast^2-\frac{\Gamma_\ast^2}{4}-\ii M_\ast\Gamma_\ast.
\]
Comparing the imaginary part with
\(\operatorname{Im}s_\ast=-\rho_0/A+O(\rho_0^2/A^2M_0^2)\) gives
\eqref{eq:qcd-linear-elastic-width-coordinate}; the real part then gives the
stated correction to \(M_\ast\).  The calculation supplies a coordinate
linearization of the pole equation.  It is appropriate for a narrow
\(\rho\)-type vector pole in an elastic window.  For a broad scalar such as
the \(f_0(500)\), the pole coordinate itself is the invariant output and a
real-axis mass-width pair is a secondary chart, if it is used at all.
\begin{remark}[How the chart is used]
For the \(\rho\), one may use the \(I=1\), \(\ell=1\) entry of
Definition~\ref{def:qcd-concrete-light-meson-pole-chart} and, when the
elastic window and narrowness assumptions are adequate, the linear pole
coordinate above.
For \(f_0(500)\), the same one-channel formula is not the correct primary
language: the broad pole is obtained by solving the analytically continued
Roy equations for \(t_0^0(s)\).  For \(\pi K\) resonances the unequal-mass
Roy--Steiner construction above fixes the kinematics and the crossing input.
For near-threshold exotic candidates, the effective-range and pole-motion
charts of Definition~\ref{def:qcd-controlled-pole-motion-chart} and the
shallow effective-range calculation above specify
what is meant by a molecular or compact coordinate.  In every case the QCD
statement is a pole or a stable mass extracted from a declared operator and
regulator setup.
\end{remark}

\subsection{Unequal-Mass and Baryon-Meson Channels}
\label{subsec:qcd-unequal-mass-baryon-meson-channels}

\begin{definition}[Unequal-mass elastic kinematics]
\label{def:qcd-unequal-mass-elastic-kinematics}
Let \(A(m_1)+B(m_2)\to A(m_1)+B(m_2)\) be elastic scattering of two stable
color-singlet hadrons, and set
\[
  s_\pm:=(m_1\pm m_2)^2,\qquad
  \lambda(s,m_1^2,m_2^2):=(s-s_+)(s-s_-).
\]
For \(s>s_+\), define the center-of-mass momentum
\[
  k_s(s):=\frac{\lambda(s,m_1^2,m_2^2)^{1/2}}{2\sqrt s}.
\]
If \(z_s\) is the cosine of the \(s\)-channel scattering angle, the
Mandelstam variables in the physical elastic region are
\begin{equation}
  t(s,z_s)=-2k_s(s)^2(1-z_s),
  \qquad
  u(s,z_s)=2(m_1^2+m_2^2)-s-t(s,z_s).
  \label{eq:qcd-unequal-mass-elastic-kinematics}
\end{equation}
The identity
\begin{equation}
  s+t+u=2(m_1^2+m_2^2)
  \label{eq:qcd-pik-mandelstam-sum}
\end{equation}
is part of the kinematic convention.  For \(\pi K\) scattering one takes
\((m_1,m_2)=(m_\pi,m_K)\).
\end{definition}

\paragraph{Elastic unequal-mass projection.}
In the kinematics of
Definition~\ref{def:qcd-unequal-mass-elastic-kinematics}, a scalar
amplitude \(T(s,t,u)\) with fixed internal quantum numbers has the
\(s\)-channel partial-wave expansion
\begin{equation}
  T(s,t(s,z_s),u(s,z_s))
  =
  16\pi
  \sum_{\ell=0}^{\infty}(2\ell+1)P_\ell(z_s)f_\ell(s),
  \label{eq:qcd-unequal-mass-partial-wave-expansion}
\end{equation}
with inverse
\begin{equation}
  f_\ell(s)
  =
  \frac1{32\pi}
  \int_{-1}^{1}P_\ell(z_s)T(s,t(s,z_s),u(s,z_s))\,\dd z_s .
  \label{eq:qcd-unequal-mass-partial-wave-projection}
\end{equation}
With
\[
  \rho_{12}(s):=\frac{2k_s(s)}{\sqrt s}
  =
  \frac{\lambda(s,m_1^2,m_2^2)^{1/2}}{s},
\]
elastic unitarity in this normalization is
\begin{equation}
  \operatorname{Im}f_\ell(s)=\rho_{12}(s)|f_\ell(s)|^2
  \label{eq:qcd-unequal-mass-elastic-unitarity}
\end{equation}
below the first inelastic threshold.
In the center-of-mass frame the incoming and outgoing spatial momenta have
magnitude \(k_s(s)\).  The energy of each species is the same before and
after scattering, hence the momentum transfer has zero energy component and
spatial magnitude squared
\[
  |\mathbf k-\mathbf k'|^2=2k_s(s)^2(1-z_s).
\]
With the Mandelstam convention used in this chapter this gives
\(t=-|\mathbf k-\mathbf k'|^2\), proving
\eqref{eq:qcd-unequal-mass-elastic-kinematics}.  The sum rule
\eqref{eq:qcd-pik-mandelstam-sum} follows directly from the definitions of
\(s,t,u\) and the four on-shell conditions.

The partial-wave inversion is the Legendre orthogonality identity
\[
  \int_{-1}^{1}P_\ell(z)P_{\ell'}(z)\,\dd z
  =
  \frac{2}{2\ell+1}\delta_{\ell\ell'}
\]
applied to \eqref{eq:qcd-unequal-mass-partial-wave-expansion}.  Finally the
two-body phase-space density in these conventions is
\(\rho_{12}(s)=2k_s(s)/\sqrt s\).  Applying \(S_\ell^\ast S_\ell=1\) to
\[
  S_\ell(s)=1+2\ii\rho_{12}(s)f_\ell(s)
\]
gives \eqref{eq:qcd-unequal-mass-elastic-unitarity}.

\begin{definition}[Roy--Steiner input]
\label{def:qcd-roy-steiner-input}
Roy--Steiner input for an unequal-mass two-hadron problem consists of:
\begin{enumerate}
  \item the masses, exact internal quantum numbers, and invariant amplitudes
  \(\mathbf T(s,t,u)\) in a declared basis of isospin or charge channels;
  \item the crossing matrices relating the chosen \(s\)-channel basis to the
  \(u\)- and \(t\)-channel bases;
  \item \(s\)-channel partial waves \(f_\ell^a(s)\) and crossed-channel
  partial waves \(g_\ell^c(t)\), each with an explicit normalization such as
  \eqref{eq:qcd-unequal-mass-partial-wave-expansion};
  \item a fixed-\(t\), fixed-\(u\), or hyperbolic dispersion domain in which
  the required subtracted Cauchy representations are valid as boundary values
  of holomorphic functions;
  \item subtraction constants, matching points, and high-energy/high-partial
  wave input sufficient to make all projected integrals converge.
\end{enumerate}
This input is a reconstruction coordinate for amplitudes.  It includes the
analyticity, crossing, subtraction, and asymptotic assumptions; QCD supplies
the channel Hilbert spaces and current or hadron matrix elements to which
those assumptions are applied.
\end{definition}

\begin{proposition}[Roy--Steiner projection construction]
\label{prop:qcd-roy-steiner-projection-construction}
Assume Roy--Steiner input in the sense of
Definition~\ref{def:qcd-roy-steiner-input}.  Let \(J_s\) and \(J_t\) be real
intervals below the first omitted inelastic thresholds in the \(s\)- and
crossed \(t\)-channels.  Then the retained \(s\)-channel partial waves have
the form
\begin{align}
  \operatorname{Re}f_\ell^a(s)
  &=
  k_\ell^a(s;\mathbf a)
  +
  \sum_{b,\ell'}
  \int_{s_+}^{\infty}\dd s'\,
  K_{\ell\ell'}^{ab}(s,s')\,\operatorname{Im}f_{\ell'}^b(s')
  \notag\\
  &\hspace{2.3cm}
  +
  \sum_{c,\ell'}
  \int_{t_{\rm th}}^{\infty}\dd t'\,
  G_{\ell\ell'}^{ac}(s,t')\,\operatorname{Im}g_{\ell'}^c(t'),
  \qquad s\in J_s,
  \label{eq:qcd-roy-steiner-s-channel-equation}
\end{align}
and the crossed-channel partial waves have analogous equations
\begin{align}
  \operatorname{Re}g_\ell^c(t)
  &=
  \widetilde k_\ell^c(t;\mathbf a)
  +
  \sum_{b,\ell'}
  \int_{s_+}^{\infty}\dd s'\,
  \widetilde G_{\ell\ell'}^{cb}(t,s')\,\operatorname{Im}f_{\ell'}^b(s')
  \notag\\
  &\hspace{2.3cm}
  +
  \sum_{d,\ell'}
  \int_{t_{\rm th}}^{\infty}\dd t'\,
  \widetilde K_{\ell\ell'}^{cd}(t,t')\,\operatorname{Im}g_{\ell'}^d(t'),
  \qquad t\in J_t.
  \label{eq:qcd-roy-steiner-t-channel-equation}
\end{align}
Here \(\mathbf a\) denotes the declared subtraction constants.  The kernels
\(K,G,\widetilde G,\widetilde K\) are the Legendre projections of the
subtracted Cauchy denominators after the crossed-channel discontinuities have
been rewritten with the declared crossing matrices.
\end{proposition}

\begin{proof}
Begin with the subtracted Cauchy representation in the chosen
Roy--Steiner domain.  Its discontinuity terms are located on the physical
cuts of the \(s\), \(u\), and \(t\) variables.  The right-hand \(s\)-channel
cut is already expressed in the basis whose partial waves are
\(f_\ell^a\).  The \(u\)- and \(t\)-channel cuts are first moved into their
physical variables and then rewritten in the chosen \(s\)-channel basis using
the crossing matrices included in the input.  This produces a fixed-angle
representation
\[
\begin{aligned}
  &\operatorname{Re}T^a(s,t(s,z_s),u(s,z_s))
  \\
  &\quad =
  P^a(s,z_s;\mathbf a)
  +\sum_{b,\ell'}\int_{s_+}^{\infty}\dd s'\,
    R_{\ell'}^{ab}(s,z_s;s')\,\operatorname{Im}f_{\ell'}^b(s')
  \\
  &\qquad\quad
  +\sum_{c,\ell'}\int_{t_{\rm th}}^{\infty}\dd t'\,
    H_{\ell'}^{ac}(s,z_s;t')\,\operatorname{Im}g_{\ell'}^c(t').
\end{aligned}
\]
The convergence part of the input justifies the interchange of the
partial-wave sums, dispersion integrals, and \(z_s\)-projection on \(J_s\).
Projecting with
\[
  \frac1{32\pi}\int_{-1}^{1}P_\ell(z_s)(\,\cdot\,)\,\dd z_s
\]
defines
\[
  k_\ell^a(s)=\frac1{32\pi}\int_{-1}^{1}P_\ell(z_s)P^a(s,z_s)\,\dd z_s,
\]
and similarly defines \(K_{\ell\ell'}^{ab}\) and
\(G_{\ell\ell'}^{ac}\) as the same projection of \(R_{\ell'}^{ab}\) and
\(H_{\ell'}^{ac}\).  This gives
\eqref{eq:qcd-roy-steiner-s-channel-equation}.  For the \(t\)-channel, use
the crossed-channel angle \(z_t\), the \(t\)-channel partial-wave
normalization, and the crossed basis of amplitudes before projection.  The
same subtracted Cauchy denominators then project to
\(\widetilde G_{\ell\ell'}^{cd}\) and
\(\widetilde K_{\ell\ell'}^{cd}\), while the polynomial part projects to
\(\widetilde k_\ell^c(t)\).  This gives
\eqref{eq:qcd-roy-steiner-t-channel-equation}.  No
additional dynamical assumption enters at the projection stage; all dynamical
input is already contained in the analytic domain, crossing matrices,
absorptive parts, and subtraction constants of the input.
\end{proof}

\subsection{Quarkonium, Baryon, and Hybrid Source Charts}
\label{subsec:qcd-quarkonium-baryon-hybrid-charts}

\begin{definition}[Quarkonium spectroscopy chart]
\label{def:qcd-quarkonium-spectroscopy-chart}
A neutral heavy quarkonium entry is a color-singlet \(Q\bar Q\) spectral
coordinate with flavor \(Q=c\) or \(Q=b\), radial label \(n\), orbital angular
momentum \(L=0,1,2,\ldots\), total heavy-quark spin \(S=0,1\), and total
angular momentum \(J\).  The spectroscopic notation is
\begin{equation}
  n\,{}^{2S+1}L_J,
  \qquad
  L=S,P,D,F,\ldots\quad\hbox{for}\quad L=0,1,2,3,\ldots .
  \label{eq:qcd-quarkonium-spectroscopic-notation}
\end{equation}
For hidden flavor, charge conjugation is an exact quantum number in the
isospin-symmetric QCD-plus-QED-decoupled convention.  The low-lying chart is:
\[
\begin{array}{c|c|c|c}
\hbox{state type} & n\,{}^{2S+1}L_J & J^{PC} & \hbox{standard entries}\\
\hline
\hbox{pseudoscalar} & n\,{}^1S_0 & 0^{-+} & \eta_c(nS),\ \eta_b(nS)\\
\hbox{vector} & n\,{}^3S_1 & 1^{--} & J/\psi,\ \psi(nS),\ \Upsilon(nS)\\
\hbox{scalar triplet} & n\,{}^3P_0 & 0^{++} & \chi_{c0}(nP),\ \chi_{b0}(nP)\\
\hbox{axial triplet} & n\,{}^3P_1 & 1^{++} & \chi_{c1}(nP),\ \chi_{b1}(nP)\\
\hbox{tensor triplet} & n\,{}^3P_2 & 2^{++} & \chi_{c2}(nP),\ \chi_{b2}(nP)\\
\hbox{singlet \(P\)-wave} & n\,{}^1P_1 & 1^{+-} & h_c(nP),\ h_b(nP)
\end{array}
\]
The entries in the last column are names for QCD spectral levels or
sheet-labeled poles in the corresponding channels.  The chart by itself is
not a potential model.  A prediction of the level \(M_{n\,{}^{2S+1}L_J}\)
requires a regulator or an effective Hamiltonian with a stated heavy-quark
mass scheme, matching coefficients, ultrasoft treatment, and error estimate.
\end{definition}

\paragraph{Quarkonium \(J^{PC}\) rule.}
For a neutral \(Q\bar Q\) color-singlet state in the chart of
Definition~\ref{def:qcd-quarkonium-spectroscopy-chart}, the allowed angular
momenta are
\[
  J\in\{|L-S|,|L-S|+1,\ldots,L+S\},
\]
and the parity and charge-conjugation quantum numbers are
\begin{equation}
  P=(-1)^{L+1},
  \qquad
  C=(-1)^{L+S}.
  \label{eq:qcd-quarkonium-pc-rule}
\end{equation}
The angular-momentum statement is the Clebsch--Gordan decomposition of the
orbital representation \(L\) with the total spin representation \(S\).  For a
fermion-antifermion pair, intrinsic parities multiply to \(-1\).  The orbital
wavefunction contributes \((-1)^L\) under spatial inversion, giving
\[
  P=(-1)(-1)^L=(-1)^{L+1}.
\]
Charge conjugation exchanges the quark and antiquark.  In the center-of-mass
frame this exchange reverses the relative coordinate, contributing
\((-1)^L\).  The spin wavefunction contributes \(+1\) on the spin-singlet
state and \(-1\) on the spin-triplet state after the fermion exchange
convention is included; this factor is \((-1)^S\) relative to the orbital
factor in the standard \(Q\bar Q\) convention.  Therefore
\[
  C=(-1)^L(-1)^S=(-1)^{L+S}.
\]
Substitution gives the entries in
Definition~\ref{def:qcd-quarkonium-spectroscopy-chart}: for example
\({}^1S_0\) has \(0^{-+}\), \({}^3S_1\) has \(1^{--}\),
\({}^3P_J\) has \(J^{++}\), and \({}^1P_1\) has \(1^{+-}\).

Heavy quarkonium uses a different controlled coordinate system.  If
\[
  m_Q\gg m_Qv\gg m_Qv^2
\]
is a useful hierarchy, the pNRQCD Hamiltonian of the preceding sections
predicts the charmonium and bottomonium level spacings through a potential
coordinate, spin-dependent \(1/m_Q\) operators, and ultrasoft corrections.
The spin-averaged levels, hyperfine splittings, and fine structure are then
matrix elements in a stated effective Hamiltonian.  They are not predictions
of a nonrelativistic potential unless the matching coefficients, mass scheme,
and ultrasoft treatment have been fixed as part of the input.  The angular
momentum algebra that turns these matched operators into centroid, fine, and
hyperfine coordinates is the finite-dimensional content of the \(J^{PC}\)
rule above and the spin algebra paragraph.

Baryon spectroscopy is likewise a QCD spectral problem, not a three-body
mechanics problem imposed from outside.  For \(N_f=3\) light flavors, the
lowest baryons organize approximately into flavor \(SU(3)\) octet and
decuplet representations when the quark masses are close.  To first order in
an octet symmetry-breaking Hamiltonian, the octet masses may be written as
\begin{equation}
  M(Y,I)=M_0+aY
  +b\left(I(I+1)-\frac{Y^2}{4}\right),
  \label{eq:qcd-baryon-octet-gmo-chart}
\end{equation}
where \(Y\) is hypercharge and \(I\) is isospin.  This coordinate chart gives
the Gell-Mann--Okubo relation
\begin{equation}
  2(M_N+M_\Xi)=3M_\Lambda+M_\Sigma .
  \label{eq:qcd-baryon-gell-mann-okubo}
\end{equation}
For the decuplet, first-order octet breaking gives equal spacing along the
hypercharge ladder.  These relations are symmetry consequences of a stated
approximation; the actual QCD masses require lattice, dispersive, and chiral
extrapolation inputs.

Assume that first-order \(SU(3)\)-breaking baryon masses are described by
\eqref{eq:qcd-baryon-octet-gmo-chart} in the octet and by
\(M_{10}(Y)=M_{10,0}+cY\) in the decuplet.  Then
\eqref{eq:qcd-baryon-gell-mann-okubo} holds and
\begin{equation}
  M_{\Sigma^\ast}-M_\Delta
  =
  M_{\Xi^\ast}-M_{\Sigma^\ast}
  =
  M_\Omega-M_{\Xi^\ast}.
  \label{eq:qcd-decuplet-equal-spacing}
\end{equation}
For the octet,
\[
  (Y,I)_N=(1,\tfrac12),\quad
  (Y,I)_\Lambda=(0,0),\quad
  (Y,I)_\Sigma=(0,1),\quad
  (Y,I)_\Xi=(-1,\tfrac12).
\]
Substitution in \eqref{eq:qcd-baryon-octet-gmo-chart} gives
\[
  M_N=M_0+a+\frac b2,\quad
  M_\Xi=M_0-a+\frac b2,\quad
  M_\Lambda=M_0,\quad
  M_\Sigma=M_0+2b,
\]
which implies \eqref{eq:qcd-baryon-gell-mann-okubo}.  For the decuplet the
hypercharges are \(1,0,-1,-2\).  A linear function of \(Y\) has equal
successive differences, giving \eqref{eq:qcd-decuplet-equal-spacing}.

In the large-\(N_c\) limit with \(N_f\) fixed, the baryon operator
\eqref{eq:qcd-baryon-local-source} contains \(N_c\) quark fields.  The
pair-counting argument around \eqref{eq:qcd-large-n-baryon-pair-counting}
gives a mass of order \(N_c\), while collective spin-flavor splittings are
typically \(O(1/N_c)\) in the rotor regime.  This is a controlled limit of QCD
under its own hypotheses, and it explains why baryons require a separate
large-\(N_c\) analysis from mesons.  It does not license importing
constituent-quark spectra as QCD predictions.

\begin{definition}[Baryon-meson resonance reconstruction data]
\label{def:qcd-baryon-meson-resonance-data}
Fix conserved quantum numbers: baryon number one, total isospin \(I\), total
strangeness \(S\), total angular momentum and parity \(J^P\), and any other
exact charges in the chosen QCD mass convention.  A two-body baryon-meson
resonance reconstruction data consist of:
\begin{enumerate}
  \item a finite list of stable color-singlet two-body channels
  \(a=(B_a,M_a)\) in the chosen energy window, with baryon mass \(M_{B_a}\),
  meson mass \(m_{M_a}\), spin \(s_{B_a}\), meson spin \(s_{M_a}\), and
  intrinsic parities \(P_{B_a},P_{M_a}\);
  \item a list of retained orbital angular momenta \(L_a\) and spin couplings
  producing the declared \(J^P\);
  \item the infinite-volume partial-wave matrix \(t_{ab}^{J^P}(s)\), with
  phase-space matrix
  \[
    \rho_a(s)=
    \frac{2k_a(s)}{\sqrt s},
    \qquad
    \sqrt s=
    \sqrt{M_{B_a}^2+k_a(s)^2}
    +
    \sqrt{m_{M_a}^2+k_a(s)^2};
  \]
  \item a reconstruction method for \(t^{J^P}\), for example a
  coupled-channel finite-volume determinant or a Roy--Steiner representation;
  \item a sheet label \(\Sigma=\operatorname{diag}(\sigma_a)\), with
  \(\sigma_a=\pm1\), specifying which channel momenta have been continued
  through their unitarity cuts;
  \item pole positions \(s_\ast\) and residue vectors \(g_a\) defined by
  \begin{equation}
    t_{ab}^{(\Sigma),J^P}(s)
    =
    \frac{g_a g_b}{s_\ast-s}
    +h_{ab}(s),
    \label{eq:qcd-baryon-resonance-residue-vector}
  \end{equation}
  where \(h_{ab}\) is holomorphic at \(s_\ast\) on the same sheet.
\end{enumerate}
The invariant content is the pole position, the sheet, and the residue
projective vector \(g_a\), together with the operator and regulator data by
which the amplitude was obtained.  Names such as radial excitation,
molecule, compact multiquark configuration, or meson-baryon bound state are
interpretive coordinates on these data.
\end{definition}

\paragraph{Pseudoscalar-meson--baryon partial-wave quantum numbers.}
Let a positive-parity spin-\(\frac12\) baryon \(B\) scatter with a
pseudoscalar meson \(M\) of spin \(0\) and parity \(-1\).  In a two-body
partial wave of orbital angular momentum \(L\), the allowed total angular
momenta and parity are
\begin{equation}
  J\in
  \begin{cases}
    \{\frac12\},&L=0,\\
    \{L-\frac12,L+\frac12\},&L\ge1,
  \end{cases}
  \qquad
  P=(-1)^{L+1}.
  \label{eq:qcd-pseudoscalar-baryon-jp-rule}
\end{equation}
Consequently an \(S\)-wave channel has \(J^P=\frac12^-\), while a
\(P\)-wave channel has \(J^P=\frac12^+\) or \(J^P=\frac32^+\).
The two-body Hilbert space at fixed center-of-mass momentum decomposes under
spatial rotations as the tensor product of the orbital representation of spin
\(L\), the baryon spin representation \(\frac12\), and the meson spin
representation \(0\).  Therefore the total spin content is
\[
  L\otimes\frac12
  =
  \begin{cases}
    \frac12,&L=0,\\
    (L-\frac12)\oplus(L+\frac12),&L\ge1,
  \end{cases}
\]
which gives the allowed \(J\)'s.  Parity multiplies the intrinsic parities by
the orbital factor \((-1)^L\):
\[
  P=P_B P_M(-1)^L=(+1)(-1)(-1)^L=(-1)^{L+1}.
\]
The \(S\)- and \(P\)-wave cases are the first two values of \(L\).

This elementary representation-theoretic step is often where baryon
resonance nomenclature enters the analysis.  The \(\Delta(1232)\) channel is
the \(I=\frac32\), \(J^P=\frac32^+\) \(P\)-wave \(\pi N\) amplitude.  A
Roper-type nucleon resonance is a pole in the \(I=\frac12\),
\(J^P=\frac12^+\) channel, with residues into the retained \(\pi N\),
\(\eta N\), and multi-hadron channels determined by the chosen coupled-channel
amplitude.  The \(\Lambda(1405)\)-type channel has \(I=0\), \(S=-1\),
\(J^P=\frac12^-\), and its minimal two-body description couples the
\(\bar K N\) and \(\pi\Sigma\) \(S\)-wave channels.  In each case the QCD
statement is the sheet-labeled pole of
Definition~\ref{def:qcd-baryon-meson-resonance-data}.  A finite-volume
calculation supplies color-singlet energy levels through finite source-family
correlators; the pole information is then reconstructed only after a
coupled-channel quantization condition and analytic continuation have been
specified.  A dispersive
calculation supplies it through the corresponding analyticity and crossing
assumptions.

\begin{proposition}[Two-channel sheet continuation and residue vector]
\label{prop:qcd-two-channel-sheet-residue}
Fix a partial wave and an energy neighborhood in which two stable
color-singlet two-body channels are the only open channels retained in the
analysis.  Suppose that, in this neighborhood, the coupled-channel amplitude
admits a \(K\)-matrix representation
\begin{equation}
  t^{\rm phys}(s)
  =
  \left[
    K(s)^{-1}
    -\ii\,\operatorname{diag}\!\bigl(\rho_1(s),\rho_2(s)\bigr)
  \right]^{-1},
  \label{eq:qcd-two-channel-physical-kmatrix}
\end{equation}
where \(K(s)^{-1}\) is a complex-analytic symmetric matrix that is real
symmetric on the physical real interval, and each \(\rho_a(s)\) is the
chosen square-root phase-space branch for channel \(a\).  For a sheet label
\(\sigma=(\sigma_1,\sigma_2)\), \(\sigma_a=\pm1\), define
\begin{equation}
  D_\sigma(s)
  :=
  K(s)^{-1}
  -\ii\,\operatorname{diag}\!\bigl(\sigma_1\rho_1(s),\sigma_2\rho_2(s)\bigr),
  \qquad
  t^{(\sigma)}(s):=D_\sigma(s)^{-1}.
  \label{eq:qcd-two-channel-sheet-denominator}
\end{equation}
Then a resonance pole on this sheet is a zero of
\begin{equation}
  \Delta_\sigma(s):=\det D_\sigma(s).
  \label{eq:qcd-two-channel-pole-determinant}
\end{equation}
If \(s_\ast\) is a simple zero of \(\Delta_\sigma\) and
\(\operatorname{rank}D_\sigma(s_\ast)=1\), then
\begin{equation}
  t^{(\sigma)}(s)
  =
  -\,\frac{\operatorname{adj}D_\sigma(s_\ast)}
          {\Delta_\sigma'(s_\ast)}\,
    \frac1{s_\ast-s}
  +O(1).
  \label{eq:qcd-two-channel-residue-adjugate}
\end{equation}
Because \(D_\sigma(s_\ast)\) is complex symmetric, the adjugate in
\eqref{eq:qcd-two-channel-residue-adjugate} has rank one and can be written
as \(c\,v_av_b\), where \(v\) spans
\(\ker D_\sigma(s_\ast)\).  Thus the projective residue vector in
Definition~\ref{def:qcd-baryon-meson-resonance-data} is the null direction of
the analytically continued denominator, with the overall complex phase fixed
only after choosing a residue normalization.

In the constant local coordinate chart
\[
  K^{-1}=
  \begin{pmatrix}a&b\\ b&c\end{pmatrix},
  \qquad
  \rho_a(s)=\rho_a,
\]
the pole equation on sheet \(\sigma\) is
\begin{equation}
  (a-\ii\sigma_1\rho_1)(c-\ii\sigma_2\rho_2)-b^2=0.
  \label{eq:qcd-two-channel-local-pole-equation}
\end{equation}
\end{proposition}

\begin{proof}
Equation~\eqref{eq:qcd-two-channel-physical-kmatrix} is the representation
assumed in the proposition.  On the physical interval it is equivalent to the
two-channel elastic-unitarity identity
\[
  \operatorname{Im}\bigl(t^{\rm phys}(s)^{-1}\bigr)
  =
  -\operatorname{diag}(\rho_1(s),\rho_2(s))
\]
on the physical interval.  Analytic continuation through the unitarity cut of
channel \(a\) changes the sign of the square root \(k_a(s)\), hence of
\(\rho_a(s)\), and leaves the other channel branches unchanged.  This gives
\eqref{eq:qcd-two-channel-sheet-denominator}.  Since the matrix inverse is
meromorphic wherever the entries are analytic, its poles are precisely the
zeros of \(\det D_\sigma\), unless the adjugate vanishes to sufficient order
to cancel the zero.  Under the stated rank-one and simple-zero hypotheses,
the adjugate does not vanish, and
\[
  D_\sigma(s)^{-1}
  =
  \frac{\operatorname{adj}D_\sigma(s)}
       {\Delta_\sigma(s)}
  =
  \frac{\operatorname{adj}D_\sigma(s_\ast)}
       {(s-s_\ast)\Delta_\sigma'(s_\ast)}
  +O(1),
\]
which is \eqref{eq:qcd-two-channel-residue-adjugate} after rewriting
\((s-s_\ast)^{-1}=-(s_\ast-s)^{-1}\).  A \(2\times2\) matrix of rank one has
rank-one adjugate; for a complex symmetric matrix the left and right null
directions agree, so this rank-one matrix is proportional to \(vv^T\).  The
constant formula \eqref{eq:qcd-two-channel-local-pole-equation} is the
determinant of the displayed \(2\times2\) denominator.
\end{proof}

This proposition is the minimal algebra behind statements about the
\(\Lambda(1405)\)-type \(\bar K N\)-\(\pi\Sigma\) sector, hidden-charm
pentaquark channels, and near-threshold heavy-meson channels.  The number of
poles, their locations, and their residues are outputs of a finite-volume or
dispersive reconstruction.  The phrase ``molecular'' can be used only after a
specific residue, pole-motion, or source-overlap coordinate has been declared;
it is not an independent QCD definition.

\begin{definition}[Light-baryon spectroscopy charts]
\label{def:qcd-light-baryon-spectroscopy-charts}
Work with baryon number \(B=1\) and three light flavors
\((u,d,s)\).  Let strangeness be denoted by \(S_{\rm str}\), so the
hypercharge used in \eqref{eq:qcd-baryon-octet-gmo-chart} is
\(Y=B+S_{\rm str}=1+S_{\rm str}\).  The ground-state flavor multiplet charts
used below are:
\begin{center}
\small
\begin{tabular}{c|c|c|c|c}
\textbf{multiplet} & \textbf{entry} & \(\mathbf I\) & \(\mathbf Y\) & \(\mathbf J^P\)\\
\hline
\(\mathbf 8\) & \(N=(p,n)\) & \(\frac12\) & \(1\) & \(\frac12^+\)\\
\(\mathbf 8\) & \(\Lambda\) & \(0\) & \(0\) & \(\frac12^+\)\\
\(\mathbf 8\) & \(\Sigma=(\Sigma^+,\Sigma^0,\Sigma^-)\) & \(1\) & \(0\) & \(\frac12^+\)\\
\(\mathbf 8\) & \(\Xi=(\Xi^0,\Xi^-)\) & \(\frac12\) & \(-1\) & \(\frac12^+\)\\
\hline
\(\mathbf {10}\) & \(\Delta\) & \(\frac32\) & \(1\) & \(\frac32^+\)\\
\(\mathbf {10}\) & \(\Sigma^\ast\) & \(1\) & \(0\) & \(\frac32^+\)\\
\(\mathbf {10}\) & \(\Xi^\ast\) & \(\frac12\) & \(-1\) & \(\frac32^+\)\\
\(\mathbf {10}\) & \(\Omega^-\) & \(0\) & \(-2\) & \(\frac32^+\)
\end{tabular}
\end{center}
The table is a chart of exact conserved quantum numbers and approximate
flavor-\(SU(3)\) organization.  The masses are QCD spectral values obtained
from the color-singlet Hamiltonian or from a regulator converging to it.
Equation~\eqref{eq:qcd-baryon-gell-mann-okubo} and
\eqref{eq:qcd-decuplet-equal-spacing} are first-order coordinates on these
spectral values; they are not definitions of the baryons.
\end{definition}

\begin{definition}[Excited light-baryon channel chart]
\label{def:qcd-excited-light-baryon-channel-chart}
An excited light-baryon entry in this monograph is specified by:
\begin{enumerate}[label=(\roman*)]
\item exact charges \(B=1\), \(I\), \(S_{\rm str}\), \(J^P\), and any
  additional exact discrete quantum numbers of the chosen mass convention;
\item a finite list of two-body and effective multi-hadron channels retained
  in the reconstruction of Definition~\ref{def:qcd-baryon-meson-resonance-data};
\item a sheet-labeled pole \(s_\ast\) of the corresponding color-singlet
  amplitude, or an isolated stable level if the state lies below every open
  channel in the stated volume and mass regime;
\item the residue projective vector into the retained channels, when the
  interpretation of the pole requires an overlap or composition coordinate.
\end{enumerate}
The following entries are the minimal light-baryon examples that must be
carried through any concrete coupled-channel analysis:
\begin{enumerate}[label=(\alph*)]
\item \(\Delta(1232)\): charges
  \((I,S_{\rm str},J^P)=(\frac32,0,\frac32^+)\); retained channel
  \(\pi N\) in the \(P\)-wave; QCD output a nearby second-sheet pole of
  \(t_{\pi N,\pi N}^{3/2^+}\).
\item \(N(1440)\)-type channel: charges
  \((\frac12,0,\frac12^+)\); retained channels
  \(\pi N,\eta N,\pi\pi N\); QCD output a coupled-channel pole whose
  radial-label interpretation comes from pole motion and source overlaps.
\item \(N(1535)\)-type channel: charges
  \((\frac12,0,\frac12^-)\); retained channels
  \(\pi N,\eta N,K\Lambda,K\Sigma\); QCD output an \(S\)-wave pole with
  residue data.
\item \(\Sigma(1385)\): charges \((1,-1,\frac32^+)\); retained channels
  \(\pi\Lambda,\pi\Sigma,\bar K N\); QCD output a strange \(P\)-wave pole.
\item \(\Lambda(1405)\)-type channel: charges \((0,-1,\frac12^-)\);
  retained channels \(\bar K N,\pi\Sigma\); QCD output one or more
  sheet-labeled \(S\)-wave poles.
\item \(\Lambda(1520)\): charges \((0,-1,\frac32^-)\); retained channels
  \(\bar K N,\pi\Sigma\); QCD output a \(D\)-wave dominated pole coordinate.
\end{enumerate}
The words ``Roper'', ``\(\Lambda(1405)\)'', and similar labels are therefore
shorthand for entries in this chart.  A lattice calculation supplies the
finite-volume spectrum and operator overlaps; a dispersive or
Hamiltonian-scattering analysis supplies an analytic continuation.  Without
that continuation there is no invariant resonance mass or width.
\end{definition}

\begin{definition}[Heavy-baryon QCD/HQET chart]
\label{def:qcd-heavy-baryon-hqet-chart}
Let \(Q=c\) or \(b\), and consider one-heavy-quark baryons \(Qqq\).  After
the heavy-quark velocity \(v^\mu\) and mass scheme for \(m_Q\) are fixed,
the color-singlet Hilbert-space sector admits an HQET expansion in which the
light degrees of freedom carry angular momentum \(j_\ell\) and flavor
quantum numbers.  The basic entries are:
\begin{center}
\small
\begin{tabular}{p{0.24\textwidth}|p{0.28\textwidth}|c|p{0.23\textwidth}}
\textbf{family} & \textbf{light flavor} & \(\mathbf{j_\ell}\) &
\textbf{\(J^P\) at leading order}\\
\hline
\(\Lambda_Q,\Xi_Q\) & \(\overline{\mathbf 3}\) of light flavor &
\(0\) & \(\frac12^+\)\\
\(\Sigma_Q,\Xi'_Q,\Omega_Q\) & \(\mathbf 6\) of light flavor &
\(1\) & \(\frac12^+,\frac32^+\)
\end{tabular}
\end{center}
The leading HQET statement is a spectral degeneracy at fixed \(j_\ell\) and
light-flavor representation, before \(1/m_Q\) chromomagnetic terms are
included.  Doubly heavy baryons such as \(\Xi_{cc}\), \(\Xi_{bc}\), and
\(\Omega_{cc}\) are QCD sectors with two heavy fields and one light field.
They may be organized by heavy-diquark variables only after a scale
separation and color-\(\overline{\mathbf 3}\) matching calculation has been
specified.
\end{definition}

\begin{definition}[Near-threshold heavy spectroscopy chart]
\label{def:qcd-near-threshold-heavy-spectroscopy-chart}
Heavy-flavor spectroscopy contains stable baryon sectors and resonance or
near-threshold pole sectors.  The following chart fixes the QCD object before
any model language is introduced:
\begin{enumerate}[label=(\alph*)]
\item \(\Xi_{cc},\Omega_{cc}\): exact charges \(B=1,\ C=2\); minimal
  sources \(ccq\); QCD output isolated color-singlet masses when the levels
  lie below the relevant strong thresholds.
\item \(\Lambda_b,\Sigma_b,\Xi_b,\Omega_b\): exact charges \(B=1\) with
  bottom flavor \(-1\); minimal sources \(bqq\); QCD output stable or narrow
  levels organized by the HQET chart above.
\item \(P_c\)-type hidden-charm baryon sector: exact charges \(B=1,\ C=0\)
  with hidden \(c\bar c\); retained channels
  \(J/\psi\,p,\Sigma_c\bar D,\Sigma_c\bar D^\ast,\ldots\); QCD output a
  sheet-labeled baryon-meson pole.
\item \(T_{cc}\)-type sector: exact charges \(B=0,\ C=2\); retained channels
  \(DD^\ast,D^\ast D^\ast,\ldots\); QCD output a near-threshold meson-meson
  pole or bound-state pole.
\item \(X(6900)\)-type sector: exact charges \(B=0,\ C=0\) with hidden
  \(c\bar c c\bar c\); retained channels
  \(J/\psi\,J/\psi,\eta_c\eta_c,\ldots\); QCD output a coupled
  heavy-quarkonium pole.
\end{enumerate}
The entries with \(B=0\) are included because they use the same pole-motion
and effective-range machinery as the baryon-number-one pentaquark channels.
Labels such as compact tetraquark, hadronic molecule, or diquark-antidiquark
are then coordinates on residues, pole motion under quark-mass variation,
large-\(N_c\) behavior, and source-overlap coordinates.  They are not primitive
spectral definitions.
\end{definition}

\begin{definition}[Glueball and hybrid spectroscopy source families]
\label{def:qcd-glueball-hybrid-source-families}
Fix \(J^{PC}\), flavor quantum numbers, and a finite-volume regulator with a
transfer matrix.  A glueball source family is a finite-dimensional subspace
\(\mathcal V_G\) spanned by gauge-invariant pure-glue operators, such as
traces of small Wilson loops or local curvature polynomials after operator
renormalization.  A flavor-singlet meson source family is a subspace
\(\mathcal V_M\) spanned by color-singlet quark bilinears and multi-hadron
operators with the same exact quantum numbers.  A hybrid source family is a
subspace \(\mathcal V_H\) spanned by gauge-invariant sources in which quark
fields are connected by Wilson lines and one or more local curvature
insertions, for example representative operator families of the form
\(\bar q(x)\Gamma\,U(x,y)F_{\mu\nu}(y)U(y,z)q(z)\) after the endpoint and
composite-operator renormalizations are specified.

The QCD spectrum is the spectrum extracted from the full correlation matrix
on the span of these source spaces and their multi-hadron completions.  The
labels glueball, hybrid, and flavor-singlet meson are source-family
coordinates on eigenvectors and pole residues.  They become invariant only
after a supplementary limit or comparison is stated, for example the pure
Yang--Mills limit, the large-\(N_c\) limit, or a declared quark-mass
deformation path.
\end{definition}

\begin{controlledapproximation}[Two-state mixing coordinate]
\label{ca:qcd-two-state-glueball-hybrid-mixing}
A two-state mixing matrix is a coordinate on a controlled finite-dimensional
projection of the transfer-matrix problem, not a definition of the QCD
spectrum.  It is admissible only after the source subspaces, the inner
product, the fitting window, and the error made by omitting other levels have
been stated.  Within such a projection, suppose two normalized basis vectors
\(|G\rangle\in\mathcal V_G\) and
\(|M\rangle\in\mathcal V_M\oplus\mathcal V_H\) give the self-adjoint matrix
\begin{equation}
  H_{\rm eff}(x)=
  \begin{pmatrix}
    M_G(x)&h(x)\\
    h(x)&M_M(x)
  \end{pmatrix},
  \label{eq:qcd-two-state-mixing-hamiltonian}
\end{equation}
where \(x\) is a declared deformation coordinate, such as a quark mass or
\(1/N_c\).
\end{controlledapproximation}

\begin{remark}[Avoided crossing in a two-state spectroscopy chart]
\label{rem:qcd-two-state-avoided-crossing}
In the controlled approximation
\ref{ca:qcd-two-state-glueball-hybrid-mixing}, the two projected energies are
\begin{equation}
  E_\pm(x)
  =
  \frac{M_G(x)+M_M(x)}2
  \pm
  \sqrt{
    \left(\frac{M_M(x)-M_G(x)}2\right)^2+h(x)^2
  }.
  \label{eq:qcd-two-state-avoided-crossing-eigenvalues}
\end{equation}
If \(M_G(x_0)=M_M(x_0)\) and \(h(x_0)\ne0\), the projected levels are
separated by \(2|h(x_0)|\).  Away from degeneracy, a real orthogonal
diagonalization angle \(\theta(x)\) may be chosen so that
\begin{equation}
  \tan 2\theta(x)=\frac{2h(x)}{M_M(x)-M_G(x)}.
  \label{eq:qcd-two-state-mixing-angle}
\end{equation}
\par
The characteristic polynomial of
\eqref{eq:qcd-two-state-mixing-hamiltonian} is
\[
  E^2-(M_G+M_M)E+M_GM_M-h^2 .
\]
Solving this quadratic gives
\eqref{eq:qcd-two-state-avoided-crossing-eigenvalues}.  At \(M_G=M_M\), the
square root is \(|h|\), so the separation is \(2|h|\).  For the angle, write
the rotated off-diagonal entry as
\[
  \frac12(M_G-M_M)\sin2\theta+h\cos2\theta .
\]
Setting it to zero gives
\((M_M-M_G)\sin2\theta=2h\cos2\theta\), which is
\eqref{eq:qcd-two-state-mixing-angle} when the denominator is nonzero.  At
degeneracy the diagonalizing angle is \(\theta=\pi/4\) modulo basis signs.
\end{remark}

The proposition explains why an eigenstate cannot be assigned an invariant
glueball or hybrid percentage from a two-by-two table alone.  The invariant
quantities are the QCD energies or pole positions and the declared matrix
elements.  The percentages are coordinates depending on the finite source
space, normalization, and deformation path.

\paragraph{Heavy-baryon spin average removes the chromomagnetic coordinate.}
Fix a one-heavy-quark baryon family with \(j_\ell>0\).  Suppose that, through
order \(1/m_Q\), the only spin-dependent term in the effective Hamiltonian
on the two-state multiplet \(J=j_\ell\pm\frac12\) is
\begin{equation}
  H_{\rm mag}
  =
  \frac{2\kappa_{j_\ell}}{m_Q}\,
  \mathbf S_Q\cdot\mathbf j_\ell ,
  \label{eq:qcd-heavy-baryon-chromomagnetic-hamiltonian}
\end{equation}
where \(\mathbf S_Q^2=\frac34\) and
\(\mathbf j_\ell^2=j_\ell(j_\ell+1)\).  Then the degeneracy-weighted
spin average
\begin{equation}
  \overline M_{j_\ell}
  :=
  \frac{(2j_\ell)M_{j_\ell-\frac12}
        +(2j_\ell+2)M_{j_\ell+\frac12}}
       {4j_\ell+2}
  \label{eq:qcd-heavy-baryon-spin-average}
\end{equation}
is independent of \(\kappa_{j_\ell}\) at this order.  In particular,
for the \(\Sigma_Q,\Sigma_Q^\ast\)-type doublet, \(j_\ell=1\) and
\begin{equation}
  \overline M_{\Sigma_Q}
  =
  \frac{M_{\Sigma_Q}+2M_{\Sigma_Q^\ast}}{3}
  \label{eq:qcd-sigma-heavy-spin-average}
\end{equation}
has no chromomagnetic \(1/m_Q\) contribution in the stated scheme.

On a state with total spin \(J\),
\[
  \mathbf S_Q\cdot\mathbf j_\ell
  =
  \frac12\left[J(J+1)-j_\ell(j_\ell+1)-\frac34\right].
\]
For \(J_-=j_\ell-\frac12\), this eigenvalue is
\[
  \frac12\left[(j_\ell-\tfrac12)(j_\ell+\tfrac12)
  -j_\ell(j_\ell+1)-\frac34\right]
  =
  -\frac{j_\ell+1}{2}.
\]
For \(J_+=j_\ell+\frac12\), it is
\[
  \frac12\left[(j_\ell+\tfrac12)(j_\ell+\tfrac32)
  -j_\ell(j_\ell+1)-\frac34\right]
  =
  \frac{j_\ell}{2}.
\]
The degeneracies are \(2J_-+1=2j_\ell\) and \(2J_++1=2j_\ell+2\).  Hence
\[
  (2j_\ell)\left(-\frac{j_\ell+1}{2}\right)
  +(2j_\ell+2)\left(\frac{j_\ell}{2}\right)=0 .
\]
The spin-average \eqref{eq:qcd-heavy-baryon-spin-average} therefore cancels
the coefficient \(\kappa_{j_\ell}\) independently of its value.  The remaining
terms are the heavy-quark mass coordinate, the spin-independent light-sector
energy, the kinetic \(1/m_Q\) coordinate, and higher-order corrections.

\begin{remark}[Large-\(N_c\) baryon catalog coordinate]
The large-\(N_c\) baryon discussion preceding
\eqref{eq:qcd-large-n-baryon-pair-counting} gives a second controlled
organization of baryon spectra.  Its invariant inputs are the QCD
Hamiltonian, baryon number one, fixed \(N_f\), and the large-\(N_c\) scaling
limit.  Under the Hartree and rotor hypotheses stated there, baryon masses
have an \(O(N_c)\) common part and \(O(1/N_c)\) spin-flavor splittings.  The
contracted \(SU(2N_f)\) algebra is a leading-order relation among baryon
matrix elements in this limit.  It complements the finite-\(N_c\) spectral
and coupled-channel charts above; it does not replace the QCD pole or
eigenvalue definitions.
\end{remark}

\subsection{Current Matrix Elements and Forward Sum Rules}
\label{subsec:qcd-current-matrix-elements-forward-sum-rules}

\begin{definition}[Nucleon current matrix elements]
\label{def:qcd-nucleon-current-matrix-elements}
Let \(|N(p,s)\rangle\) be an isolated one-nucleon state in the color-singlet
QCD Hilbert space, with \(p^2=-M_N^2\), \(p^0>0\), and spin label \(s\).  In
the mostly-plus spinor convention of Section~\ref{sec:spinor-conventions},
use the same Dirac--Pauli basis as in
Definition~\ref{def:qed-electron-electromagnetic-form-factors}.  For the Hermitian
electromagnetic current \(J_{\rm em}^\mu\), normalized by the flavor charges
of the quarks, define the nucleon Dirac and Pauli form factors by
\begin{equation}
  \langle N(p',s')|J_{\rm em}^\mu(0)|N(p,s)\rangle
  =
  \frac{\ii e}{(2\pi)^3}\,
  \bar u_{s'}(p')
  \left[
    \gamma^\mu F_1^N(k^2)
    +
    \frac1{M_N}S^{\mu\nu}k_\nu F_2^N(k^2)
  \right]
  u_s(p),
  \qquad k:=p-p',
  \label{eq:qcd-nucleon-dirac-pauli-form-factors}
\end{equation}
where \(e>0\) is the elementary charge and
\(S^{\mu\nu}=-\ii[\gamma^\mu,\gamma^\nu]/4\).  The charge normalization is
\[
  F_1^p(0)=1,\qquad F_1^n(0)=0.
\]
The anomalous magnetic moment coordinate is \(F_2^N(0)\), and the magnetic
moment in nuclear magnetons is \(F_1^N(0)+F_2^N(0)\) after the conventional
normalization by \(e/(2M_N)\).

For spacelike transfer set
\[
  Q^2:=-k^2>0,\qquad
  \tau:=\frac{Q^2}{4M_N^2}.
\]
The Sachs coordinates are the linear combinations
\begin{equation}
  G_E^N(Q^2)=F_1^N(-Q^2)-\tau F_2^N(-Q^2),
  \qquad
  G_M^N(Q^2)=F_1^N(-Q^2)+F_2^N(-Q^2).
  \label{eq:qcd-nucleon-sachs-form-factors}
\end{equation}
They are useful Breit-frame coordinates for elastic electron--nucleon
scattering, but the nonperturbative definition is the current matrix element
\eqref{eq:qcd-nucleon-dirac-pauli-form-factors}.  A lattice, dispersive, or
experimental extraction must state the current renormalization, finite-volume
matrix-element method, excited-state removal, and analytic continuation or
fit ansatz used to obtain the functions of \(Q^2\).
\end{definition}

\begin{remark}[Sachs and Dirac--Pauli coordinates]
\label{rem:qcd-nucleon-sachs-coordinate-transform}
For \(\tau\ne -1\), the coordinate change
\eqref{eq:qcd-nucleon-sachs-form-factors} is invertible:
\begin{equation}
  F_1^N(-Q^2)
  =
  \frac{G_E^N(Q^2)+\tau G_M^N(Q^2)}{1+\tau},
  \qquad
  F_2^N(-Q^2)
  =
  \frac{G_M^N(Q^2)-G_E^N(Q^2)}{1+\tau}.
  \label{eq:qcd-nucleon-sachs-inverse}
\end{equation}
At \(Q^2=0\),
\[
  G_E^N(0)=F_1^N(0),\qquad
  G_M^N(0)=F_1^N(0)+F_2^N(0).
\]
\par
The two equations in \eqref{eq:qcd-nucleon-sachs-form-factors} form the
linear system
\[
  \begin{pmatrix}G_E\\G_M\end{pmatrix}
  =
  \begin{pmatrix}1&-\tau\\ 1&1\end{pmatrix}
  \begin{pmatrix}F_1\\F_2\end{pmatrix}.
\]
The determinant is \(1+\tau\).  Inverting the matrix gives
\eqref{eq:qcd-nucleon-sachs-inverse}.  Setting \(Q^2=0\), hence \(\tau=0\),
gives the stated zero-transfer normalizations.  No model of nucleon
substructure enters this argument; it is only a change of coordinates on the
same current matrix element.
\end{remark}

\begin{definition}[Isovector axial form factors]
\label{def:qcd-nucleon-axial-form-factors}
Let \(A_a^\mu\) be the renormalized isovector axial current, with
\(a=1,2,3\) and flavor generator \(\tau_a/2\).  Between one-nucleon states,
parity, Lorentz covariance, and isospin give the decomposition
\begin{equation}
  \langle N(p',s')|A_a^\mu(0)|N(p,s)\rangle
  =
  \frac{\ii}{(2\pi)^3}\,
  \bar u_{s'}(p')
  \left[
    \gamma^\mu\gamma_5\,G_A(Q^2)
    +
    \frac{k^\mu}{2M_N}\gamma_5\,G_P(Q^2)
  \right]
  \frac{\tau_a}{2}\,
  u_s(p),
  \label{eq:qcd-nucleon-axial-form-factors}
\end{equation}
again with \(Q^2=-k^2\).  The number \(g_A:=G_A(0)\) is the axial charge in
this current normalization.  The induced pseudoscalar form factor \(G_P\)
contains the pion-pole contribution dictated by the chiral Ward identity and
the pion effective theory of Chapter~\ref{ch:volume-ii-21}; beyond that pole,
its regular part is another QCD matrix element.  Thus nucleon axial data are
part of the same ab initio program as the mass spectrum: they are not
defined by a quark-model wave function but by renormalized current insertions
between color-singlet Hilbert-space states.
\end{definition}

\begin{proposition}[Stable-current three-point extraction]
\label{prop:qcd-stable-current-three-point-extraction}
Let a reflection-positive finite-volume regulator of QCD on a spatial torus
\(\mathbb T_L^3\) be fixed, with Hilbert space \(\mathcal H_L\), vacuum
\(\Omega_L\), self-adjoint Hamiltonian \(H_L\ge0\), and momentum operators
\(\mathbf P_L\).  Let \(J_{\mathcal R}\) be a renormalized, gauge-invariant
current in a specified operator-renormalization scheme \(\mathcal R\), and let
\(J_{\mathcal R,\mathbf q}\) denote its spatial Fourier component carrying
momentum \(\mathbf q\).  Choose momenta
\(\mathbf p,\mathbf p'\in(2\pi/L)\mathbb Z^3\) with
\(\mathbf p'=\mathbf p+\mathbf q\).  Suppose that in the two relevant
baryon-number-one momentum sectors there are normalized one-nucleon eigenvectors
\[
  H_L|N,L;\mathbf p,r\rangle=E_L(\mathbf p)|N,L;\mathbf p,r\rangle,\qquad
  H_L|N,L;\mathbf p',r'\rangle=E_L(\mathbf p')|N,L;\mathbf p',r'\rangle,
\]
isolated from the rest of the spectrum in the same quantum-number channels by
positive gaps \(\Delta_L(\mathbf p)\) and \(\Delta_L(\mathbf p')\).  Here
\(r,r'\) label chosen rows in the relevant finite-volume symmetry irreducible
representations; after the \(L\to\infty\) limit these rows are matched to the
usual spin labels.  Let \(\mathcal O_{\mathbf p,r}^\dagger\) and
\(\mathcal O_{\mathbf p',r'}^\dagger\) be gauge-invariant baryon sources,
projected to the indicated momentum and row, and set
\[
  \psi_{\mathbf p,r}:=\mathcal O_{\mathbf p,r}^\dagger\Omega_L,\qquad
  \psi_{\mathbf p',r'}:=\mathcal O_{\mathbf p',r'}^\dagger\Omega_L.
\]
Assume their one-particle overlaps
\[
  Z_{\mathbf p,r}:=\langle N,L;\mathbf p,r|\psi_{\mathbf p,r}\rangle,\qquad
  Z_{\mathbf p',r'}:=
  \langle N,L;\mathbf p',r'|\psi_{\mathbf p',r'}\rangle
\]
are nonzero.  At fixed regulator, assume \(J_{\mathcal R,\mathbf q}\) is
bounded on the finite-energy subspace generated by these sources; on a lattice
with compact gauge variables and local fermion bilinears this is the ordinary
boundedness statement for the regulated local insertion.

Define the Euclidean two- and three-point functions
\begin{align}
  C_2(t;\mathbf p,r)
  &:=
  \langle\psi_{\mathbf p,r},e^{-H_Lt}\psi_{\mathbf p,r}\rangle,
  \label{eq:qcd-two-point-transfer-matrix-form}\\
  C_3^{J_{\mathcal R}}(t,\tau;\mathbf p',r';\mathbf p,r)
  &:=
  \langle
    \psi_{\mathbf p',r'},
    e^{-H_L(t-\tau)}J_{\mathcal R,\mathbf q}e^{-H_L\tau}
    \psi_{\mathbf p,r}
  \rangle,
  \qquad 0<\tau<t .
  \label{eq:qcd-three-point-transfer-matrix-form}
\end{align}
Then
\begin{equation}
  C_2(t;\mathbf p,r)
  =
  |Z_{\mathbf p,r}|^2e^{-E_L(\mathbf p)t}
  +R_2(t;\mathbf p,r),
  \qquad
  |R_2(t;\mathbf p,r)|
  \le
  \|\eta_{\mathbf p,r}\|^2
  e^{-(E_L(\mathbf p)+\Delta_L(\mathbf p))t},
  \label{eq:qcd-two-point-gap-estimate}
\end{equation}
where
\(\eta_{\mathbf p,r}:=(1-P_{\mathbf p,r})\psi_{\mathbf p,r}\) and
\(P_{\mathbf p,r}\) is the spectral projection onto
\(|N,L;\mathbf p,r\rangle\).  Moreover
\begin{align}
  C_3^{J_{\mathcal R}}(t,\tau;\mathbf p',r';\mathbf p,r)
  &=
  \overline{Z_{\mathbf p',r'}}Z_{\mathbf p,r}\,
  e^{-E_L(\mathbf p')(t-\tau)}e^{-E_L(\mathbf p)\tau}
  \langle N,L;\mathbf p',r'|
    J_{\mathcal R,\mathbf q}
  |N,L;\mathbf p,r\rangle
  \nonumber\\
  &\hspace{1cm}
  +R_3(t,\tau;\mathbf p',r';\mathbf p,r),
  \label{eq:qcd-three-point-gap-expansion}
\end{align}
with the explicit bound
\begin{align}
  |R_3|
  \le
  \|J_{\mathcal R,\mathbf q}\|
  \Big(&
    \|\eta_{\mathbf p',r'}\|\,|Z_{\mathbf p,r}|\,
    e^{-(E_L(\mathbf p')+\Delta_L(\mathbf p'))(t-\tau)}
    e^{-E_L(\mathbf p)\tau}
  \nonumber\\
  &+
    |Z_{\mathbf p',r'}|\,\|\eta_{\mathbf p,r}\|\,
    e^{-E_L(\mathbf p')(t-\tau)}
    e^{-(E_L(\mathbf p)+\Delta_L(\mathbf p))\tau}
  \nonumber\\
  &+
    \|\eta_{\mathbf p',r'}\|\,\|\eta_{\mathbf p,r}\|\,
    e^{-(E_L(\mathbf p')+\Delta_L(\mathbf p'))(t-\tau)}
    e^{-(E_L(\mathbf p)+\Delta_L(\mathbf p))\tau}
  \Big).
  \label{eq:qcd-three-point-gap-bound}
\end{align}
Consequently the ratio
\begin{equation}
  \mathcal R_{J_{\mathcal R}}(t,\tau)
  :=
  \frac{
    C_3^{J_{\mathcal R}}(t,\tau;\mathbf p',r';\mathbf p,r)}
  {
    \overline{Z_{\mathbf p',r'}}Z_{\mathbf p,r}\,
    e^{-E_L(\mathbf p')(t-\tau)}e^{-E_L(\mathbf p)\tau}}
  \label{eq:qcd-three-point-matrix-element-ratio}
\end{equation}
satisfies
\[
  \mathcal R_{J_{\mathcal R}}(t,\tau)
  =
  \langle N,L;\mathbf p',r'|
    J_{\mathcal R,\mathbf q}
  |N,L;\mathbf p,r\rangle
  +O(e^{-\Delta_L(\mathbf p')(t-\tau)})
  +O(e^{-\Delta_L(\mathbf p)\tau})
\]
as \(\tau\to\infty\) and \(t-\tau\to\infty\), with constants fixed by the
source norms, overlaps, and regulated-current norm above.
\end{proposition}

\begin{proof}
The proof is the spectral theorem for the finite-volume transfer-matrix
Hamiltonian.  We give the details because this is the mathematical content
behind the phrase ``extracting a form factor''.  Decompose the source vectors
into their one-particle components and orthogonal remainders,
\[
  \psi_{\mathbf p,r}
  =
  Z_{\mathbf p,r}|N,L;\mathbf p,r\rangle+\eta_{\mathbf p,r},
  \qquad
  \psi_{\mathbf p',r'}
  =
  Z_{\mathbf p',r'}|N,L;\mathbf p',r'\rangle+\eta_{\mathbf p',r'}.
\]
By the assumed spectral gaps, the spectral support of
\(\eta_{\mathbf p,r}\) lies in
\([E_L(\mathbf p)+\Delta_L(\mathbf p),\infty)\), and similarly for
\(\eta_{\mathbf p',r'}\).  Hence
\[
  \|e^{-H_Ls}\eta_{\mathbf p,r}\|
  \le
  e^{-(E_L(\mathbf p)+\Delta_L(\mathbf p))s}\|\eta_{\mathbf p,r}\|,
  \qquad s\ge0,
\]
and the corresponding estimate holds in the primed channel.  Substituting the
decomposition of \(\psi_{\mathbf p,r}\) into
\eqref{eq:qcd-two-point-transfer-matrix-form} gives the leading term
\(|Z_{\mathbf p,r}|^2e^{-E_L(\mathbf p)t}\).  The cross terms vanish because
\(\eta_{\mathbf p,r}\) is orthogonal to the one-particle eigenspace and
\(e^{-H_Lt}\) preserves the spectral decomposition.  The remaining
\(\eta\)-term is bounded by the preceding estimate, proving
\eqref{eq:qcd-two-point-gap-estimate}.

For the three-point function, insert the same decompositions on the sink and
source sides of \eqref{eq:qcd-three-point-transfer-matrix-form}.  The term in
which both spectral projections are one-particle projections is exactly the
first line of \eqref{eq:qcd-three-point-gap-expansion}.  The three remaining
terms contain an excited component on the sink side, on the source side, or on
both sides.  Applying the gap estimate to the side on which the excited
component propagates, and applying Cauchy's inequality together with the
regulated operator norm of \(J_{\mathcal R,\mathbf q}\), gives the three
terms in \eqref{eq:qcd-three-point-gap-bound}.  Division by the nonzero
leading overlap factor gives \eqref{eq:qcd-three-point-matrix-element-ratio}
and the stated asymptotic error.  Nothing in the argument assumes a constituent
description of the nucleon; the only inputs are reflection positivity,
spectral isolation, a renormalized gauge-invariant current insertion, and
nonzero overlap of the chosen source with the state.
\end{proof}

The continuum nucleon form factors of
Definition~\ref{def:qcd-nucleon-current-matrix-elements} are obtained
from the finite-volume matrix element in
Proposition~\ref{prop:qcd-stable-current-three-point-extraction} only after
the current-renormalization scheme has been fixed and the regulator and
volume limits have been taken with the chosen state normalization.  A plateau
in a ratio of Euclidean correlators is therefore evidence for the hypotheses
of the proposition in a numerical calculation; it is not itself the definition
of the form factor.

\begin{definition}[Forward Compton amplitude and polarizabilities]
\label{def:qcd-forward-compton-polarizabilities}
Let \(N\) be a stable proton or neutron state of mass \(M_N\) and spin
\(\frac12\).  Work in the nucleon rest frame and let a real photon have energy
\(\omega>0\), direction \(\mathbf n\in S^2\), and transverse polarization
representatives \(\boldsymbol\epsilon,\boldsymbol\epsilon'\) satisfying
\(\mathbf n\cdot\boldsymbol\epsilon
=\mathbf n\cdot\boldsymbol\epsilon'=0\).  The forward real-Compton amplitude
is the on-shell, connected, LSZ-reduced matrix element of two insertions of
the renormalized electromagnetic current \(J_{\rm em}^\mu\), with the photon
coupling and the one-nucleon state normalization fixed by the scattering
chapters.  For a spin state with unit spin vector \(\mathbf s\), define the
scalar forward amplitudes \(f_N(\omega)\) and \(g_N(\omega)\) by
\begin{equation}
  \mathcal A_{\gamma N\to\gamma N}^{\rm fwd}
  =
  \boldsymbol\epsilon'^{\,*}\!\cdot\boldsymbol\epsilon\, f_N(\omega)
  +
  \ii\,\mathbf s\cdot
  \bigl(\boldsymbol\epsilon'^{\,*}\times\boldsymbol\epsilon\bigr)\,
  g_N(\omega).
  \label{eq:qcd-forward-compton-fg-definition}
\end{equation}
This equation is a definition of the normalization of \(f_N,g_N\).  Let
\[
  f_N^{\rm nb}(\omega):=f_N(\omega)-f_N^{\rm Born}(\omega)
\]
denote the non-Born amplitude after subtracting the one-nucleon pole
contribution, whose normalization is specified in the low-energy expansion
below.  With this normalization the optical-theorem convention for the
absorptive non-Born amplitude is
\begin{equation}
  \operatorname{Im}f_N^{\rm nb}(\omega+\ii0)
  =
  \frac{\omega}{4\pi}\,
  \sigma_{\gamma N}^{\rm abs}(\omega),
  \qquad
  \omega\ge\omega_{\rm th},
  \label{eq:qcd-forward-compton-optical-theorem}
\end{equation}
where \(\omega_{\rm th}\) is the first inelastic threshold and
\(\sigma_{\gamma N}^{\rm abs}\) is the total real-photon absorption cross
section in the same convention, not the elastic Compton cross section.  The
spin-dependent optical theorem similarly uses the helicity-difference
absorption cross section
\(\Delta\sigma_{\gamma N}:=\sigma_{1/2}-\sigma_{3/2}\):
\begin{equation}
  \operatorname{Im}g_N(\omega+\ii0)
  =
  \frac{\omega}{8\pi}\,
  \Delta\sigma_{\gamma N}(\omega).
  \label{eq:qcd-forward-spin-compton-optical-theorem}
\end{equation}

The spin-independent electric-plus-magnetic polarizability is the coefficient
of the first non-Born term in the low-energy expansion
\begin{equation}
  f_N(\omega)
  =
  f_N^{\rm Born}(\omega)
  +
  \bigl(\alpha_E^N+\beta_M^N\bigr)\omega^2
  +O(\omega^4).
  \label{eq:qcd-forward-polarizability-definition}
\end{equation}
The Born amplitude is the one-nucleon pole contribution determined by the
charge \(F_1^N(0)\), magnetic moment \(F_1^N(0)+F_2^N(0)\), and the same
Compton Ward identity used in the QED external-state chapter.  Equation
\eqref{eq:qcd-forward-polarizability-definition} defines only the sum
\(\alpha_E^N+\beta_M^N\).  Separating \(\alpha_E^N\) from \(\beta_M^N\)
requires either non-forward Compton amplitudes or a background-field
definition with separately specified electric and magnetic external fields.
\end{definition}

\begin{proposition}[Baldin sum rule as a forward-dispersion theorem]
\label{prop:qcd-baldin-forward-dispersion}
Assume that the spin-independent forward Compton amplitude
\(f_N^{\rm nb}(\omega)\) is an even function of \(\omega\) analytic in the
complex \(\omega\)-plane cut along
\((-\infty,-\omega_{\rm th}]\cup[\omega_{\rm th},\infty)\), real on the
real interval \(|\omega|<\omega_{\rm th}\), and sufficiently bounded at
infinity so that the once-subtracted Cauchy formula below has no boundary
term.  Then
\begin{equation}
  f_N^{\rm nb}(\omega)-f_N^{\rm nb}(0)
  =
  \frac{2\omega^2}{\pi}
  \int_{\omega_{\rm th}}^\infty
  \frac{\operatorname{Im}f_N^{\rm nb}(\omega'+\ii0)}
       {\omega'(\omega'^2-\omega^2)}
  \,\dd\omega',
  \qquad |\omega|<\omega_{\rm th}.
  \label{eq:qcd-forward-compton-subtracted-dispersion}
\end{equation}
If the small-\(\omega\) expansion
\eqref{eq:qcd-forward-polarizability-definition} exists and the integral may
be expanded at \(\omega=0\), then
\begin{equation}
  \alpha_E^N+\beta_M^N
  =
  \frac{1}{2\pi^2}
  \int_{\omega_{\rm th}}^\infty
  \frac{\sigma_{\gamma N}^{\rm abs}(\omega')}{\omega'^2}\,\dd\omega' .
  \label{eq:qcd-baldin-sum-rule}
\end{equation}
\end{proposition}

\begin{proof}
Set \(F(\omega):=f_N^{\rm nb}(\omega)-f_N^{\rm nb}(0)\).  By crossing in the
forward channel \(F\) is even.  Apply Cauchy's theorem to
\(F(z)/(z^2-\omega^2)\) on a contour surrounding the two cuts and use the
assumed large-\(|z|\) bound to drop the circle at infinity.  The discontinuity
on the positive cut is
\[
  F(\omega'+\ii0)-F(\omega'-\ii0)
  =
  2\ii\,\operatorname{Im}f_N^{\rm nb}(\omega'+\ii0).
\]
The negative cut gives the same contribution after the change
\(\omega'\mapsto-\omega'\), because \(F\) is even.  The residues at
\(z=\pm\omega\) then give
\eqref{eq:qcd-forward-compton-subtracted-dispersion}.  For
\(|\omega|<\omega_{\rm th}\),
\[
  \frac{1}{\omega'^2-\omega^2}
  =
  \frac{1}{\omega'^2}
  +O(\omega^2),
\]
under the stated integrability assumption.  Therefore the coefficient of
\(\omega^2\) in \(F(\omega)\) is
\[
  \frac{2}{\pi}
  \int_{\omega_{\rm th}}^\infty
  \frac{\operatorname{Im}f_N^{\rm nb}(\omega'+\ii0)}{\omega'^3}\,\dd\omega' .
\]
Using the optical theorem
\eqref{eq:qcd-forward-compton-optical-theorem} gives
\eqref{eq:qcd-baldin-sum-rule}.  The proof shows precisely where the analytic
and high-energy hypotheses enter; QCD supplies the current matrix elements,
while the dispersion relation is an additional analyticity statement about the
real-photon forward amplitude.
\end{proof}

\begin{definition}[Forward spin polarizabilities]
\label{def:qcd-forward-spin-polarizabilities}
Keep the normalization of the spin-dependent forward Compton amplitude
\(g_N\) in \eqref{eq:qcd-forward-compton-fg-definition} and the helicity
difference
\[
  \Delta\sigma_{\gamma N}(\omega)
  :=
  \sigma_{1/2}(\omega)-\sigma_{3/2}(\omega),
\]
where the subscript is the total \(\gamma N\) helicity in the rest frame of
the target.  Let \(F_1^N(q^2)\) and \(F_2^N(q^2)\) be the Dirac and Pauli form
factors in the current matrix element
\eqref{eq:qcd-nucleon-dirac-pauli-form-factors}, and set
\[
  Q_N:=F_1^N(0),\qquad
  \kappa_N:=F_2^N(0).
\]
Thus \(Q_p=1\), \(Q_n=0\), and \(\kappa_N\) is the anomalous magnetic moment
in units \(e/(2M_N)\) in the conventions of this chapter.  The
spin-dependent low-energy theorem fixes the term linear in \(\omega\):
\begin{equation}
  g_N(\omega)
  =
  -\,\frac{e^2\kappa_N^2}{8\pi M_N^2}\,\omega
  +\gamma_0^N\,\omega^3
  +O(\omega^5).
  \label{eq:qcd-forward-spin-low-energy-definition}
\end{equation}
Equation \eqref{eq:qcd-forward-spin-low-energy-definition} defines the
forward spin polarizability \(\gamma_0^N\) once the one-nucleon Born
normalization has been fixed by \(F_1^N(0)\) and \(F_2^N(0)\).  The absence of
a \(Q_N^2\) term in the coefficient of \(\omega\) is part of the low-energy
Ward-identity statement: the point-charge contribution is spin independent in
the forward limit, while the anomalous magnetic moment controls the leading
spin-dependent Born amplitude.
\end{definition}

\begin{proposition}[Gerasimov--Drell--Hearn and forward spin-polarizability
sum rules]
\label{prop:qcd-gdh-forward-spin-dispersion}
Assume that \(g_N(\omega)\), after the Compton Ward identity has combined the
\(s\)-channel, \(u\)-channel, and local Born pieces into a function regular at
\(\omega=0\), is an odd function of \(\omega\) analytic in the complex
\(\omega\)-plane cut along
\((-\infty,-\omega_{\rm th}]\cup[\omega_{\rm th},\infty)\), real on
\(|\omega|<\omega_{\rm th}\), and sufficiently bounded at infinity that the
unsubtracted odd dispersion relation below has no boundary term.  Assume also
that the threshold and high-energy behavior allow the first two terms of the
small-\(\omega\) expansion of the dispersion integral.  Then, in the
normalization of \eqref{eq:qcd-forward-spin-compton-optical-theorem},
\begin{equation}
  g_N(\omega)
  =
  \frac{2\omega}{\pi}
  \int_{\omega_{\rm th}}^\infty
  \frac{\operatorname{Im}g_N(\omega'+\ii0)}
       {\omega'^2-\omega^2}\,\dd\omega',
  \qquad |\omega|<\omega_{\rm th},
  \label{eq:qcd-forward-spin-odd-dispersion}
\end{equation}
and therefore
\begin{align}
  \int_{\omega_{\rm th}}^\infty
  \frac{\Delta\sigma_{\gamma N}(\omega')}{\omega'}\,\dd\omega'
  &=
  -\,\frac{\pi e^2}{2M_N^2}\,\kappa_N^2
  =
  -\,\frac{2\pi^2\alpha_{\rm em}}{M_N^2}\,\kappa_N^2,
  \label{eq:qcd-gdh-sum-rule}\\
  \gamma_0^N
  &=
  \frac{1}{4\pi^2}
  \int_{\omega_{\rm th}}^\infty
  \frac{\Delta\sigma_{\gamma N}(\omega')}{\omega'^3}\,\dd\omega' .
  \label{eq:qcd-forward-spin-polarizability-sum-rule}
\end{align}
Here \(\alpha_{\rm em}:=e^2/(4\pi)\).  The sign in
\eqref{eq:qcd-gdh-sum-rule} follows from the convention
\(\Delta\sigma_{\gamma N}=\sigma_{1/2}-\sigma_{3/2}\).
\end{proposition}

\begin{proof}
Let \(\omega\) lie in the cut plane with \(|\omega|<\omega_{\rm th}\).  Apply
Cauchy's theorem to \(g_N(z)/(z^2-\omega^2)\) on a contour encircling the two
cuts and small circles around \(z=\pm\omega\).  The residue contribution is
\[
  \frac{g_N(\omega)}{2\omega}
  +
  \frac{g_N(-\omega)}{-2\omega}
  =
  \frac{g_N(\omega)}{\omega},
\]
because \(g_N\) is odd.  The large circle gives no contribution by
hypothesis.  On the positive cut the discontinuity is
\[
  g_N(\omega'+\ii0)-g_N(\omega'-\ii0)
  =
  2\ii\,\operatorname{Im}g_N(\omega'+\ii0).
\]
On the negative cut, the change of variable \(\omega'\mapsto-\omega'\) gives
the same contribution, because oddness of \(g_N\) makes
\(\operatorname{Im}g_N(-\omega'+\ii0)=
\operatorname{Im}g_N(\omega'+\ii0)\) with the opposite orientation of the
cut.  Combining the two cuts gives
\eqref{eq:qcd-forward-spin-odd-dispersion}.

For \(|\omega|<\omega_{\rm th}\),
\[
  \frac{1}{\omega'^2-\omega^2}
  =
  \frac{1}{\omega'^2}
  +
  \frac{\omega^2}{\omega'^4}
  +O(\omega^4)
\]
under the stated integrability assumption.  Hence
\[
  g_N(\omega)
  =
  \omega\,\frac{2}{\pi}
  \int_{\omega_{\rm th}}^\infty
  \frac{\operatorname{Im}g_N(\omega'+\ii0)}{\omega'^2}\,\dd\omega'
  +
  \omega^3\,\frac{2}{\pi}
  \int_{\omega_{\rm th}}^\infty
  \frac{\operatorname{Im}g_N(\omega'+\ii0)}{\omega'^4}\,\dd\omega'
  +O(\omega^5).
\]
Substitution of
\eqref{eq:qcd-forward-spin-compton-optical-theorem} turns the coefficient of
\(\omega\) into
\[
  \frac{1}{4\pi^2}
  \int_{\omega_{\rm th}}^\infty
  \frac{\Delta\sigma_{\gamma N}(\omega')}{\omega'}\,\dd\omega',
\]
and the coefficient of \(\omega^3\) into the right side of
\eqref{eq:qcd-forward-spin-polarizability-sum-rule}.  Comparing the linear
coefficient with the low-energy theorem
\eqref{eq:qcd-forward-spin-low-energy-definition} gives
\[
  \int_{\omega_{\rm th}}^\infty
  \frac{\Delta\sigma_{\gamma N}(\omega')}{\omega'}\,\dd\omega'
  =
  4\pi^2\left(-\,\frac{e^2\kappa_N^2}{8\pi M_N^2}\right),
\]
which is \eqref{eq:qcd-gdh-sum-rule}.  Thus the sum rule is not an independent
model assumption about nucleon structure; it is the equality of two
definitions of the same coefficient, one from the low-energy current Ward
identity and one from the absorptive spectral representation, under the stated
analyticity and boundedness hypotheses.
\end{proof}

The practical hierarchy of spectroscopic control is therefore:
\begin{itemize}
  \item Stable mesons and baryons are isolated color-singlet mass eigenvalues;
  their controlled extraction uses the finite-volume transfer matrix and
  continuum extrapolation.
  \item Elastic resonances are second-sheet poles of partial waves; their
  controlled extraction uses Luscher, Roy, or Roy--Steiner reconstruction.
  \item Heavy quarkonium is a color-singlet \(Q\bar Q\) spectrum; its
  controlled extraction uses a matched NRQCD or pNRQCD Hamiltonian and
  lattice checks.
  \item Hybrids and glueballs are operator-overlap sectors and pole data; their
  controlled extraction uses declared color-singlet source families,
  finite-volume energies or scattering amplitudes, large-\(N_c\) scaling, and
  mixing analysis.
  \item Exotic near-threshold states are poles plus residues in coupled
  channels; their controlled extraction uses finite-volume coupled-channel
  amplitudes.
\end{itemize}
Every entry in this list is a statement about QCD observables.  The words
``quark model'', ``molecule'', ``diquark'', and ``flux-tube excitation'' may
summarize patterns in a chosen coordinate system, but the monograph treats
them as derived interpretations only after the corresponding spectral or pole
input has been defined.

\section{Euclidean Current Sum Rules and Borel Windows}
\label{sec:qcd-current-sum-rules-borel-windows}

QCD current sum rules are linear identities for gauge-invariant current
correlators.  Their mathematical core consists of a spectral representation, a
spacelike OPE, and a chosen linear functional that suppresses subtraction
constants and weights different parts of the spectrum.  A finite-parameter
spectral ansatz is an additional approximation layer, whose status is fixed by
the error estimates stated below.

Let \(J(x)\) be a renormalized local color-singlet operator with fixed
quantum numbers, for example \(J=\bar q\Gamma q\),
\(\operatorname{tr}F_{\mu\nu}F^{\mu\nu}\), or a baryonic color singlet.  In
Euclidean momentum notation \(Q^2>0\), define the connected two-point
function
\begin{equation}
  \Pi_J(Q^2;\mu)
  :=
  \int_{\mathbb R^4}\dd^4x\,
  \ee^{\ii Q\cdot x}
  \langle J(x)J^\dagger(0)\rangle_{\rm conn,\mu},
  \qquad Q^2:=Q_\mu Q_\mu .
  \label{eq:qcd-euclidean-current-correlator}
\end{equation}
When \(J\) carries Lorentz indices, \(\Pi_J\) denotes one scalar invariant
after projecting onto a specified tensor structure.  Positivity of the
Euclidean correlator for \(J^\dagger J\) gives a positive spectral measure in
channels where reflection positivity applies.  With enough subtractions, the
spacelike correlator has the form
\begin{equation}
  \Pi_J(Q^2;\mu)
  =
  P_N(Q^2;\mu)
  +
  \int_{s_{\rm th}}^\infty
  \frac{\rho_J(s;\mu)}{s+Q^2}\,\dd s ,
  \label{eq:qcd-current-dispersion-subtracted}
\end{equation}
where \(P_N\) is a polynomial fixed by the subtraction convention,
\(s_{\rm th}\) is the lowest physical threshold in the channel, and
\(\rho_J\) is the spectral density associated with the same current
normalization.  Equation~\eqref{eq:qcd-current-dispersion-subtracted} is a
dispersion representation of a color-singlet Wightman distribution.

At large spacelike \(Q^2\), the current product has an operator-product
expansion
\begin{equation}
  \Pi_J(Q^2;\mu)
  \sim
  \sum_a C_{J,a}(Q^2,\mu)\,
  \langle \mathcal O_a(0;\mu)\rangle ,
  \qquad Q^2\to+\infty ,
  \label{eq:qcd-current-correlator-ope}
\end{equation}
with local gauge-invariant operators \(\mathcal O_a\) of increasing dimension
and Wilson coefficients \(C_{J,a}\).  The symbol \(\sim\) means an asymptotic
expansion in a specified sector of the spacelike \(Q^2\)-plane, after the
renormalization scheme and operator basis have been fixed.  Vacuum
expectation values such as \(\langle\bar q q\rangle_\mu\) or
\(\langle\operatorname{tr}F^2\rangle_\mu\) are scheme-dependent coordinates of
the vacuum state.  They become physical only through the full renormalized
combination in a specified observable.

\begin{definition}[Borel--Laplace sum-rule functional]
\label{def:qcd-borel-laplace-sum-rule-functional}
For a function \(F(Q^2)\) whose large-\(Q^2\) behavior permits the limit, set
\begin{equation}
  \mathcal B_{M^2}[F]
  :=
  \lim_{\substack{n,Q^2\to\infty\\ Q^2/n=M^2}}
  \frac{(Q^2)^n}{(n-1)!}
  \left(-\frac{\dd}{\dd Q^2}\right)^n F(Q^2),
  \qquad M^2>0 .
  \label{eq:qcd-borel-transform-definition}
\end{equation}
The positive parameter \(M^2\) is the Borel or Laplace scale.  The associated
sum-rule coordinate is
\begin{equation}
  \mathcal S_J(M^2;\mu)
  :=
  \mathcal B_{M^2}\Pi_J
  =
  \frac1{M^2}\int_{s_{\rm th}}^\infty
  \ee^{-s/M^2}\rho_J(s;\mu)\,\dd s ,
  \label{eq:qcd-borel-spectral-coordinate}
\end{equation}
provided the Borel transform can be interchanged with the spectral integral.
\end{definition}

\paragraph{Borel transform of the subtracted dispersion kernel.}
The functional in Definition~\ref{def:qcd-borel-laplace-sum-rule-functional}
annihilates every polynomial in \(Q^2\).  For \(s\ge0\),
\begin{equation}
  \mathcal B_{M^2}\!\left[\frac1{s+Q^2}\right]
  =
  \frac1{M^2}\ee^{-s/M^2}.
  \label{eq:qcd-borel-kernel-result}
\end{equation}
For an integer \(m\ge1\),
\begin{equation}
  \mathcal B_{M^2}\!\left[(Q^2)^{-m}\right]
  =
  \frac{1}{(m-1)!\,(M^2)^m}.
  \label{eq:qcd-borel-inverse-power-result}
\end{equation}
If \(F(Q^2)\) is a polynomial of degree \(d\), then
\((-\partial_{Q^2})^nF=0\) for \(n>d\), so
\(\mathcal B_{M^2}F=0\).  For the dispersion kernel,
\[
  \left(-\frac{\dd}{\dd Q^2}\right)^n\frac1{s+Q^2}
  =
  \frac{n!}{(s+Q^2)^{n+1}}.
\]
Putting \(Q^2=nM^2\) gives
\[
  \frac{(Q^2)^n}{(n-1)!}
  \frac{n!}{(s+Q^2)^{n+1}}
  =
  \frac1{M^2}
  \left(1+\frac{s}{nM^2}\right)^{-n-1}
  \longrightarrow
  \frac1{M^2}\ee^{-s/M^2}.
\]
For \((Q^2)^{-m}\),
\[
  \left(-\frac{\dd}{\dd Q^2}\right)^n(Q^2)^{-m}
  =
  \frac{\Gamma(m+n)}{\Gamma(m)}(Q^2)^{-m-n}.
\]
The Borel expression is therefore
\[
  \frac{\Gamma(m+n)}{\Gamma(m)\Gamma(n)}
  (nM^2)^{-m}.
\]
Since \(\Gamma(n+m)/\Gamma(n)=n^m(1+O(n^{-1}))\) for fixed integer \(m\), the
limit is \eqref{eq:qcd-borel-inverse-power-result}.

\begin{controlledapproximation}[QCD Borel sum-rule window]
\label{ca:qcd-borel-sum-rule-window}
A Borel sum-rule analysis in a channel \(J\) consists of:
\begin{enumerate}
  \item a renormalized current \(J\), tensor projection, and scheme at scale
  \(\mu\);
  \item a truncated OPE
  \(\Pi_J^{(D_{\max})}(Q^2;\mu)\) with a stated estimate for the omitted
  operator dimensions and perturbative orders after applying
  \(\mathcal B_{M^2}\);
  \item a spectral ansatz for \(\rho_J\), often a low-energy part with finitely
  many parameters plus a perturbative continuum above a threshold \(s_0\);
  \item a nonempty interval \(M^2\in[M_{\min}^2,M_{\max}^2]\) on which the
  omitted-OPE estimate and the spectral-tail estimate are both controlled.
\end{enumerate}
Only the equality
\eqref{eq:qcd-borel-spectral-coordinate} with the exact \(\rho_J\) is a
consequence of the dispersion representation and the Borel transform.  The
finite-parameter extraction of a mass, residue, or coupling is a controlled
approximation only after the spectral ansatz and the two error estimates above
have been supplied.
\end{controlledapproximation}

For a single isolated stable state in the channel, the spectral measure has an
atom \(Z_R\delta(s-m_R^2)\).  If the remaining spectral weight is represented
by a continuum model \(\rho_{\rm cont}(s)\), then
\eqref{eq:qcd-borel-spectral-coordinate} becomes
\begin{equation}
  \mathcal S_J(M^2)
  =
  \frac{Z_R}{M^2}\ee^{-m_R^2/M^2}
  +
  \frac1{M^2}\int_{s_{\rm cont}}^\infty
  \ee^{-s/M^2}\rho_{\rm cont}(s)\,\dd s .
  \label{eq:qcd-single-pole-plus-continuum-borel}
\end{equation}
The common pole-plus-continuum model sets
\(\rho_{\rm cont}(s)\) equal to a perturbative spectral density above a
fit threshold \(s_0\).  Its validity is an empirical and analytic control
question about the exact spectral measure, judged by stability under \(M^2\),
\(s_0\), current choice, OPE truncation, and comparison with regulator
observables such as lattice spectra.

For a finite retained spectral model
\[
  \mathcal S(\tau)=\sum_{r=1}^N Z_r e^{-\tau m_r^2},
  \qquad
  Z_r>0,\quad \tau:=1/M^2 .
\]
the logarithmic derivative used as a mass estimator is
\begin{equation}
  -\frac{\dd}{\dd\tau}\log\mathcal S(\tau)
  =
  \frac{\sum_r Z_r m_r^2 e^{-\tau m_r^2}}
       {\sum_r Z_r e^{-\tau m_r^2}} .
  \label{eq:qcd-borel-effective-mass-weighted-average}
\end{equation}
because the numerator is \(-\mathcal S'(\tau)\).  The right side is a
normalized average with positive weights \(Z_r e^{-\tau m_r^2}\).  It equals
\(m_R^2\) for a single retained atom and otherwise averages the squared masses
present in the retained spectral model.

Thus a Borel plateau diagnoses the weighted average in
\eqref{eq:qcd-borel-effective-mass-weighted-average} changes slowly over the
chosen interval after the declared continuum subtraction and OPE truncation.
Single-state dominance is a further spectral statement about the retained
model.  The object being computed remains the current correlator
\eqref{eq:qcd-euclidean-current-correlator}; the extracted spectral parameters
inherit all assumptions of the Borel window.

\begin{controlledapproximation}[Duality and stability gate for SVZ extractions]
\label{ca:qcd-svz-duality-stability-gate}
Set \(\tau=1/M^2\).  After choosing a continuum threshold \(s_0\), the
continuum-subtracted OPE moments used to quote a hadron mass are
\begin{equation}
  \mathcal L_k^{(D_{\max})}(\tau,s_0)
  =
  \int_{s_{\rm th}}^{s_0}
  s^k \ee^{-\tau s}\rho_J^{\rm pert}(s;\mu)\,\dd s
  +
  \mathcal C_k^{(D_{\max})}(\tau,\mu),
  \qquad k=0,1,\dots ,
  \label{eq:qcd-svz-continuum-subtracted-moments}
\end{equation}
where \(\mathcal C_k^{(D_{\max})}\) denotes the Borel-transformed power
corrections retained in the declared OPE truncation.  The quoted effective
mass coordinate is
\begin{equation}
  m_{\rm eff}^2(\tau,s_0)
  :=
  \frac{\mathcal L_1^{(D_{\max})}(\tau,s_0)}
       {\mathcal L_0^{(D_{\max})}(\tau,s_0)} .
  \label{eq:qcd-svz-effective-mass-coordinate}
\end{equation}
A nonempty SVZ window is accepted only after the following finite controls are
declared on the same rectangle
\((\tau,s_0)\in I_\tau\times I_{s_0}\):
\begin{enumerate}
  \item \(\mathcal L_0^{(D_{\max})}>0\), together with a stated bound for any
  loss of positivity caused by truncating a positive exact spectral measure;
  \item an OPE-tail bound for the omitted operator dimensions and perturbative
  orders in the Borel-transformed moments;
  \item a continuum-duality remainder comparing the exact spectral weight above
  and near \(s_0\) with the perturbative continuum model in the same moments;
  \item a threshold-sensitivity budget for the allowed variation of \(s_0\);
  \item a plateau-slope budget in \(\tau\), interpreted as a spectral-width
  test rather than as a proof of a pole by itself.
\end{enumerate}
For the last two entries the finite formulas are part of the control.  If the
power terms are independent of \(s_0\), then
\begin{equation}
  \partial_{s_0}m_{\rm eff}^2
  =
  \frac{\ee^{-\tau s_0}\rho_J^{\rm pert}(s_0;\mu)}
       {\mathcal L_0^{(D_{\max})}(\tau,s_0)}
  \left(s_0-m_{\rm eff}^2(\tau,s_0)\right).
  \label{eq:qcd-svz-threshold-sensitivity}
\end{equation}
Thus a plateau obtained only by retuning \(s_0\) is not a stable mass
extraction; it has used the continuum threshold as an additional dynamical
fit coordinate.  If the retained moment is the Borel transform of a positive
spectral measure, with
\(\mathcal L_k=\int s^k e^{-\tau s}\dd\mu(s)\), then
\begin{equation}
  -\partial_\tau
  \left(\frac{\mathcal L_1}{\mathcal L_0}\right)
  =
  \frac{\mathcal L_2\mathcal L_0-\mathcal L_1^2}{\mathcal L_0^2}
  =
  \operatorname{Var}_{\tau}(s)
  \ge 0 .
  \label{eq:qcd-svz-plateau-variance}
\end{equation}
Flatness therefore measures the variance of the retained spectral distribution,
plus the declared OPE, continuum, threshold, and regulator remainders.  It is
not an independent argument for single-state dominance.

The pole-plus-continuum language has a precise finite form.  Suppose the exact
low-energy moment in the window can be decomposed as
\begin{equation}
  \mathcal L_0^{\rm exact}
  =
  F_H\ee^{-\tau m_H^2}+R_0,
  \qquad
  \mathcal L_1^{\rm exact}
  =
  F_H m_H^2\ee^{-\tau m_H^2}+R_1 ,
  \label{eq:qcd-svz-pole-remainder-moments}
\end{equation}
with \(F_H>0\).  If
\(F_H\ee^{-\tau m_H^2}>|R_0|\), then the exact logarithmic estimator obeys
\begin{equation}
  \left|
  \frac{\mathcal L_1^{\rm exact}}{\mathcal L_0^{\rm exact}}
  -m_H^2
  \right|
  \le
  \frac{|R_1|+m_H^2|R_0|}
       {F_H\ee^{-\tau m_H^2}-|R_0|}.
  \label{eq:qcd-svz-pole-remainder-bound}
\end{equation}
Equation~\eqref{eq:qcd-svz-pole-remainder-bound} is the finite version of the
usual ``ground-state dominance'' sentence.  Without such a remainder bound, a
small visual slope in \(M^2\) is only a property of the chosen ansatz and
threshold, not a controlled QCD statement about the hadron spectrum.
\end{controlledapproximation}

\begin{controlledapproximation}[Two-scale SVZ extraction window]
\label{ca:qcd-svz-two-scale-extraction-window}
The preceding gate becomes an extraction, rather than a set of local
diagnostics, only after the error components are propagated to the quoted
spectral coordinate.  Let
\(\Delta_k^{\rm win}(\tau,s_0)\) denote the aggregate difference between the
exact low-energy moment and the retained continuum-subtracted OPE moment:
\begin{equation}
  \widehat{\mathcal L}_k(\tau,s_0)
  =
  \mathcal L_k^{(D_{\max})}(\tau,s_0)
  +
  \Delta_k^{\rm win}(\tau,s_0),
  \qquad k=0,1 .
  \label{eq:qcd-svz-window-moment-residuals}
\end{equation}
Here \(\Delta_k^{\rm win}\) includes the omitted OPE tail, perturbative
truncation, continuum-duality error near the threshold, regulator or finite
volume residue, and any positivity defect introduced by using a truncated
coordinate for a positive spectral measure.  If
\(\mathcal L_0^{(D_{\max})}>|\Delta_0^{\rm win}|\), then the exact and
retained mass estimators obey
\begin{equation}
  \left|
  \frac{\widehat{\mathcal L}_1}{\widehat{\mathcal L}_0}
  -
  \frac{\mathcal L_1^{(D_{\max})}}{\mathcal L_0^{(D_{\max})}}
  \right|
  \le
  \frac{
    |\Delta_1^{\rm win}|
    +
    |m_{\rm eff}^2|\,|\Delta_0^{\rm win}|
  }{
    \mathcal L_0^{(D_{\max})}-|\Delta_0^{\rm win}|
  } .
  \label{eq:qcd-svz-window-mass-residual-bound}
\end{equation}
Indeed, subtracting the two quotients gives
\((\Delta_1^{\rm win}-m_{\rm eff}^2\Delta_0^{\rm win})/
(\mathcal L_0^{(D_{\max})}+\Delta_0^{\rm win})\).  Thus the \(k=0\)
moment error is as important as the \(k=1\) moment error: changing the
normalization denominator moves the logarithmic mass even when the first
moment error vanishes.

Choose a reference squared scale \(\Lambda_J^2\) for the channel and local
half-widths \(h_\tau,h_s\) inside the proposed rectangle
\(I_\tau\times I_{s_0}\).  A finite two-scale window score may be taken to be
\begin{equation}
\begin{aligned}
  \mathfrak E_{\rm SVZ}
  :=
  \sup_{I_\tau\times I_{s_0}}
  \Bigg[
  &\frac{
    |\Delta_1^{\rm win}|
    +
    |m_{\rm eff}^2|\,|\Delta_0^{\rm win}|
  }{
    \bigl(\mathcal L_0^{(D_{\max})}-|\Delta_0^{\rm win}|\bigr)\Lambda_J^2
  }
\\
  &+
  \frac{h_\tau|\partial_\tau m_{\rm eff}^2|}{\Lambda_J^2}
  +
  \frac{h_s|\partial_{s_0}m_{\rm eff}^2|}{\Lambda_J^2}
  +
  \eta_{\rm pos}
  \Bigg],
\end{aligned}
  \label{eq:qcd-svz-two-scale-window-score}
\end{equation}
where \(\eta_{\rm pos}\) records any normalized loss of positivity in the
retained moment.  The exact choice of norm is part of the analysis, but it
must have this structure: the OPE and continuum remainders are propagated
through the quotient, and the two independent window directions are tested
separately.  A quoted mass coordinate is then a pair consisting of a central
point \((\tau_\ast,s_{0,\ast})\) and an uncertainty budget dominated by
\(\Lambda_J^2\mathfrak E_{\rm SVZ}\), not just the value read from a graph.

This two-scale requirement rules out a common false plateau.  If the threshold
is allowed to move with \(\tau\), then along a chosen graph \(s_0=s_0(\tau)\)
\begin{equation}
  \frac{\dd}{\dd\tau}
  m_{\rm eff}^2(\tau,s_0(\tau))
  =
  \partial_\tau m_{\rm eff}^2
  +
  s_0'(\tau)\,\partial_{s_0}m_{\rm eff}^2 .
  \label{eq:qcd-svz-retuned-threshold-total-derivative}
\end{equation}
Whenever \(\partial_{s_0}m_{\rm eff}^2\ne0\), one can locally choose
\(s_0'=-(\partial_\tau m_{\rm eff}^2)/
(\partial_{s_0}m_{\rm eff}^2)\) and make the displayed derivative vanish.
Flatness along that retuned path is therefore not an independent physical
stability criterion.  It is acceptable only if the allowed threshold motion is
fixed by external spectral information and its contribution is included in
\eqref{eq:qcd-svz-two-scale-window-score}; otherwise the continuum threshold
has been used to absorb the same spectral information the sum rule was meant
to extract.
\end{controlledapproximation}

\section{Detector Observables and Energy Correlators}

A nonperturbative formulation of high-energy QCD final states begins with
detector measurements in the physical Hilbert space.  The calorimetric
observable is built from the renormalized, gauge-invariant stress tensor
\(T^{\mu\nu}\), whose normalization is fixed by the translation generators:
\[
  P^\nu=\int_{\R^3}\dd^3x\,T^{0\nu}(t,\mathbf x)
\]
on the common domain of finite-energy states where the integral is defined as
a quadratic form.

\begin{construction}[Energy-flow detector from the stress tensor]
\label{cons:qcd-energy-flow-detector}
For a real test function \(f\in C^\infty(S^2)\), define the formal
null-infinity flux quadratic form by
\begin{equation}
  \widehat{\mathcal E}(f)
  =
  \mathrm{w}\!-\!\lim_{R\to\infty}
  R^2
  \int_{-\infty}^{\infty}\dd t
  \int_{S^2}\dd\Omega(\mathbf n)\,
  f(\mathbf n)\,
  n_i\,\widehat T^{0i}(t,R\mathbf n).
  \label{eq:energy-flow-operator-smeared}
\end{equation}
The notation means that the limit is first tested between finite-energy
physical states.  The operation is part of the detector definition only after
the domain and closure properties below have been supplied.
\end{construction}

\begin{hypothesis}[Detector existence and self-adjointness]
\label{hyp:qcd-energy-flow-detector}
There is a dense scattering domain
\(\mathcal D_{\rm det}\subset\Hilb_{\rm phys}\), stable under finite products
of the smeared detector forms used below, such that for every real
\(f\in C^\infty(S^2)\) the weak limit
\eqref{eq:energy-flow-operator-smeared} exists on
\(\mathcal D_{\rm det}\), is symmetric there, and has a unique self-adjoint
closure.  For each pair \(\Psi,\Phi\in\mathcal D_{\rm det}\), the map
\[
  f\longmapsto
  \bra{\Psi}\widehat{\mathcal E}(f)\ket{\Phi}
\]
is a distribution of finite order on \(S^2\).  The constant test function
satisfies
\[
  \widehat{\mathcal E}(1)=H_{\mathrm{out}}
\]
on outgoing scattering states, where \(H_{\mathrm{out}}\) is the outgoing
Hamiltonian restricted to the physical scattering Hilbert space.  Finally,
for every angular patch \(U\subset S^2\), the quadratic form obtained by
testing \(\widehat{\mathcal E}(f)\) with \(\operatorname{supp}f\subset U\)
is the \(R\to\infty\) limit of a family of complete null-generator energy
integrals with nonnegative transverse smearings whenever \(f\ge0\).  This
last clause is the geometric detector/ANEC bridge; without it, ANEC and
calorimetric positivity are statements about different limiting operators.
\end{hypothesis}

\paragraph{Detector positivity from ANEC.}
Assume Hypothesis~\ref{hyp:qcd-energy-flow-detector}.  Assume in addition
that the stress tensor satisfies the averaged null energy condition as a
quadratic-form statement on the same scattering domain: for every future null
direction \(n^\mu\), every nonnegative transverse smearing
\(\varphi\), and every \(\Psi\in\mathcal D_{\rm det}\),
\[
  \int\dd^2x_\perp\,\varphi(x_\perp)
  \int_{-\infty}^{\infty}\dd\lambda\,
  \bra{\Psi}T_{\mu\nu}(x_\perp+\lambda n)n^\mu n^\nu\ket{\Psi}
  \geq0 .
\]
Then the self-adjoint closure of \(\widehat{\mathcal E}(f)\) is positive for
every \(f\geq0\).

Write a nonnegative smooth angular test function as a finite sum of functions
supported in coordinate patches on \(S^2\), using a partition of unity.  In
each patch, the large-\(R\) detector limit is the conformal or asymptotic
image of a null-energy integral along the complete generator labelled by
\(\mathbf n\), with transverse smearing determined by the angular support of
\(f\).  The Jacobian relating the angular measure to transverse measure is
positive.  The ANEC inequality therefore gives nonnegative quadratic form
values for each patch.  Summing the patch contributions gives
\[
  \bra{\Psi}\widehat{\mathcal E}(f)\ket{\Psi}\geq0
  \qquad (\Psi\in\mathcal D_{\rm det},\ f\geq0).
\]
Since the operator has a unique self-adjoint closure by
Hypothesis~\ref{hyp:qcd-energy-flow-detector}, positivity of the quadratic
form on a core implies positivity of the closure by the spectral theorem.

\begin{remark}[Status in QCD and in conformal theories]
\label{rem:qcd-energy-flow-status}
For free massless fields, the limit in
\eqref{eq:energy-flow-operator-smeared} is obtained by inserting the mode
expansion and applying stationary phase at null infinity.  For conformal
field theories satisfying the light-transform hypotheses of
Chapter~\ref{ch:cft-light-ray-energy-correlators}, the detector is the
stress-tensor light-ray operator and positivity is supplied by the ANEC
theorem stated there.  In four-dimensional QCD, the detector is the expected
physical calorimeter observable, but its nonperturbative construction depends
on the existence of the continuum theory, the physical scattering space, and
the hadronic asymptotic completeness problem stated in
Open Problem~\ref{op:qcd-asymptotic-completeness}.  Perturbative KLN
cancellation proves finiteness of regulator-expanded coefficients for
IRC-continuous measurements after the relevant sums over degenerate states
and the required factorization steps have been made; it does not construct
the self-adjoint detector operator on the exact physical Hilbert space.
\end{remark}

\begin{openproblem}[Energy-flow operators in four-dimensional QCD]
\label{op:qcd-energy-flow-operator-construction}
Construct the operators \(\widehat{\mathcal E}(f)\) of
Construction~\ref{cons:qcd-energy-flow-detector} in continuum
four-dimensional QCD.  The construction should start from a nonperturbative
definition of the physical Hilbert space, identify a dense finite-energy
scattering domain, prove the null-infinity weak limits, prove essential
self-adjointness or specify the self-adjoint extension selected by detector
physics, prove positivity from an ANEC theorem in the same framework, and
show that \(\widehat{\mathcal E}(1)\) agrees with the outgoing Hamiltonian on
the hadronic scattering subspace.
\end{openproblem}

Here \(\mathbf n\in S^2\) is a unit spatial direction and \(\dd\Omega\) is the
rotation-invariant measure on \(S^2\).  The distribution
\(\widehat{\mathcal E}(\mathbf n)\) is defined by
\(\widehat{\mathcal E}(f)=\int_{S^2}\dd\Omega(\mathbf n)
f(\mathbf n)\widehat{\mathcal E}(\mathbf n)\).  On an outgoing state of
stable hadrons with momenta \(p_r\), species labels \(a_r\), and the
normalization conventions of scattering theory,
\begin{equation}
  \widehat{\mathcal E}(f)
  \ket{p_1,a_1;\ldots;p_N,a_N}_{\mathrm{out}}
  =
  \left(\sum_{r=1}^N p_r^0\,f(\widehat{\mathbf p}_r)\right)
  \ket{p_1,a_1;\ldots;p_N,a_N}_{\mathrm{out}},
  \qquad
  \widehat{\mathbf p}_r:=\frac{\mathbf p_r}{|\mathbf p_r|}.
  \label{eq:energy-flow-on-hadron-state}
\end{equation}
Equation~\eqref{eq:energy-flow-on-hadron-state} is the spectral action of the
detector limit on physical hadronic asymptotic states.

\paragraph{Asymptotic multiplication model.}
The preceding formula has a precise operator-theoretic meaning once an
outgoing scattering representation has been chosen.  Let
\(\mathcal H_{\mathrm{out}}\) be the physical outgoing scattering subspace and
write it as a direct integral over hadronic scattering configurations,
\[
  \mathcal H_{\mathrm{out}}
  \simeq
  \int_{\mathcal X_{\mathrm{out}}}^{\oplus}\mathcal K_Y\,\dd\nu(Y).
\]
Here \(Y\) specifies the number of stable hadrons, their species labels,
on-shell momenta, and the discrete or continuous degeneracy labels left after
the standard Bose--Fermi symmetrizations; \(\mathcal K_Y\) is the
corresponding finite or countable multiplicity space.  The outgoing
Hamiltonian is multiplication by
\[
  E_{\mathrm{tot}}(Y)=\sum_{r\in Y}p_r^0 .
\]
For a bounded Borel function \(f:S^2\to\mathbb C\), define
\begin{equation}
  m_f(Y):=\sum_{r\in Y}p_r^0 f(\widehat{\mathbf p}_r),
  \qquad
  \mathcal E_{\mathrm{mult}}(f):=M_{m_f},
  \label{eq:energy-detector-multiplication-symbol}
\end{equation}
where \(M_{m_f}\) is the direct-integral multiplication operator with maximal
domain
\[
  \operatorname{Dom} M_{m_f}
  =
  \left\{\Psi:
  \int_{\mathcal X_{\mathrm{out}}}
  |m_f(Y)|^2\|\Psi(Y)\|_{\mathcal K_Y}^2\,\dd\nu(Y)<\infty
  \right\}.
\]
If \(f\) is real, \(M_{m_f}\) is self-adjoint; if \(f\geq0\), it is positive;
and
\begin{equation}
  |m_f(Y)|\leq \|f\|_\infty E_{\mathrm{tot}}(Y),
  \qquad
  \operatorname{Dom} H_{\mathrm{out}}
  \subset
  \operatorname{Dom} \mathcal E_{\mathrm{mult}}(f).
  \label{eq:energy-detector-hamiltonian-bound}
\end{equation}
Moreover the family is strongly commuting, because all members are
multiplication operators on the same direct-integral space.  For bounded
\(f_1,\ldots,f_k\),
\begin{equation}
  \mathcal E_{\mathrm{mult}}(f_1)\cdots
  \mathcal E_{\mathrm{mult}}(f_k)
  =
  M_{\prod_{j=1}^k m_{f_j}},
  \qquad
  \operatorname{Dom} H_{\mathrm{out}}^k
  \subset
  \operatorname{Dom}
  \bigl(\mathcal E_{\mathrm{mult}}(f_1)\cdots
  \mathcal E_{\mathrm{mult}}(f_k)\bigr).
  \label{eq:energy-detector-product-domain}
\end{equation}
Indeed
\[
  \left|\prod_{j=1}^k m_{f_j}(Y)\right|
  \leq
  \left(\prod_{j=1}^k\|f_j\|_\infty\right)
  E_{\mathrm{tot}}(Y)^k ,
\]
which gives the displayed domain inclusion.  The stress-tensor flux operator
\(\widehat{\mathcal E}(f)\) of
Construction~\ref{cons:qcd-energy-flow-detector}, when it exists, is required
to agree with \(\mathcal E_{\mathrm{mult}}(f)\) after the outgoing wave
operator identifies finite-energy scattering states with their asymptotic
configurations.  This agreement is part of the nonperturbative detector
construction problem; perturbation theory may test its matrix elements after
the detector operator has been defined.

The corresponding eventwise calorimetric measure of the final state \(X\) is
\[
  \mu_X
  =
  \sum_{r\in X}p_r^0\,\delta_{\widehat{\mathbf p}_r},
  \qquad
  \widehat{\mathcal E}_X(f)=\int_{S^2}f\,\dd\mu_X .
\]
The detector observable depends on \(X\) only through the positive measure
\(\mu_X\).  The multiplication model also fixes the meaning of coincident
detectors.  For a finite configuration \(X\),
\begin{equation}
  \prod_{j=1}^k\widehat{\mathcal E}_X(f_j)
  =
  \int_{(S^2)^k}
  f_1(\mathbf n_1)\cdots f_k(\mathbf n_k)\,
  \dd\mu_X(\mathbf n_1)\cdots \dd\mu_X(\mathbf n_k).
  \label{eq:energy-detector-product-measure}
\end{equation}
Thus the product measure includes the diagonal strata
\(\mathbf n_i=\mathbf n_j\).  A separated-angle correlator is the restriction
of the resulting distribution away from these diagonals; deleting the
diagonal weight defines a different observable and must be declared as an
additional measurement prescription.

Given a normalized incoming physical state \(\Psi_{\mathrm{in}}\), or a state
prepared by a local color-singlet current and then evolved to future infinity,
the \(k\)-point energy correlator is the distribution on \((S^2)^k\)
\begin{equation}
  \mathcal G_k(f_1,\ldots,f_k;\Psi)
  =
  {}_{\mathrm{out}}\!\bra{\Psi}
  \widehat{\mathcal E}(f_1)\cdots
  \widehat{\mathcal E}(f_k)
  \ket{\Psi}_{\mathrm{out}} .
  \label{eq:energy-correlator-definition}
\end{equation}
After inserting a complete set of physical hadronic out-states, this becomes
\begin{equation}
  \mathcal G_k(f_1,\ldots,f_k;\Psi)
  =
  \sum_X
  |\braket{X,\mathrm{out}}{\Psi,\mathrm{out}}|^2
  \prod_{\ell=1}^k
  \left(
    \sum_{r\in X} p_r^0\,f_\ell(\widehat{\mathbf p}_r)
  \right),
  \label{eq:energy-correlator-hadron-sum}
\end{equation}
with the usual replacement of the displayed sum by direct integrals over
continuous multiparticle phase space.  This formula is nonperturbative: all
states in \(X\) are physical hadronic states, and the operator being measured
is built from \(T^{\mu\nu}\).

\paragraph{Connected cumulants and contact strata.}
Let \(\mathbb P_\Psi\) denote the probability measure on outgoing
configurations induced by the normalized state \(\Psi\) in the multiplication
model.  For bounded detector tests \(f_1,\ldots,f_k\), put
\[
  X_j(Y):=\mathcal E_Y(f_j)
  =
  \int_{S^2} f_j\,\dd\mu_Y .
\]
Then \(\mathcal G_k(f_1,\ldots,f_k;\Psi)\) is the raw moment
\[
  \mathbb E_{\mathbb P_\Psi}\!\left[X_1\cdots X_k\right].
\]
The connected energy correlator in this event-ensemble sense is the cumulant
\begin{equation}
  \mathcal G_k^{\rm conn}(f_1,\ldots,f_k;\Psi)
  =
  \sum_{\pi\in\Pi_k}
  (-1)^{|\pi|-1}(|\pi|-1)!
  \prod_{B\in\pi}
  \mathcal G_{|B|}\bigl((f_i)_{i\in B};\Psi\bigr),
  \label{eq:qcd-energy-correlator-cumulant}
\end{equation}
where \(\Pi_k\) is the set of set partitions of
\(\{1,\ldots,k\}\).  Thus, for example,
\[
  \mathcal G_2^{\rm conn}(f,g)
  =
  \mathcal G_2(f,g)-\mathcal G_1(f)\mathcal G_1(g),
\]
and \(\mathcal G_3^{\rm conn}\) is obtained by subtracting the three
one-block--two-block products and adding twice the product of the three
one-point functions.  This is a subtraction of event-to-event factorization
in the probability distribution of final states.

It is a different operation from removing coincident-detector contacts.  For
a fixed event \(Y\), the product measure \(\mu_Y^{\otimes k}\) decomposes
into strata labeled by coincidences among detector labels.  Already for three
detectors,
\begin{equation}
\begin{aligned}
  \mathcal E_Y(f)\mathcal E_Y(g)\mathcal E_Y(h)
  &=
  \sum_{\substack{r,s,t\in Y\\ r,s,t\ {\rm distinct}}}
    E_rE_sE_t f_r g_s h_t                                      \\
  &\quad+
  \sum_{\substack{r,t\in Y\\ r\neq t}}
    E_r^2E_t f_r g_r h_t
  +
  \sum_{\substack{r,s\in Y\\ r\neq s}}
    E_r^2E_s f_r g_s h_r                                      \\
  &\quad+
  \sum_{\substack{r,s\in Y\\ r\neq s}}
    E_rE_s^2 f_r g_s h_s
  +
  \sum_{r\in Y}E_r^3 f_r g_r h_r ,
\end{aligned}
\label{eq:qcd-three-detector-contact-strata}
\end{equation}
where \(E_r=p_r^0\) and \(f_r=f(\widehat{\mathbf p}_r)\), etc.  The last
four terms are diagonal contact strata of the detector product; they are
present before the average over events is taken.  If
\(\mathbb P_\Psi\) is concentrated on a single event, all ensemble cumulants
of order \(k\geq2\) vanish, but the diagonal terms in
\eqref{eq:qcd-three-detector-contact-strata} remain in the raw detector
product.  Conversely, deleting diagonal strata produces a separated-detector
measurement prescription, not an ensemble-connected correlator.  A calculation
that quotes a ``connected'' or ``contact-subtracted'' energy correlator must
therefore state which of these two operations is being performed.

\paragraph{Soft and collinear continuity of smeared energy flow.}
Let \(f\in C^1(S^2)\).  Suppose a final-state calorimetric measure
\(\mu_X\) is changed either by adding one particle of energy \(\varepsilon\) at
direction \(\mathbf m\), or by replacing two particles of energies \(E_a,E_b\)
and directions \(\mathbf n_a,\mathbf n_b\) by one particle of energy
\(E_a+E_b\) in the direction
\[
  \mathbf n_{ab}
  =
  \frac{E_a\mathbf n_a+E_b\mathbf n_b}
       {|E_a\mathbf n_a+E_b\mathbf n_b|}.
\]
The resulting detector change obeys
\[
  \left|\Delta\widehat{\mathcal E}_X(f)\right|
  \leq
  \varepsilon \|f\|_\infty
\]
in the soft case, and
\[
  \left|\Delta\widehat{\mathcal E}_X(f)\right|
  \leq
  (E_a+E_b)\operatorname{Lip}(f)\,
  \operatorname{diam}\{\mathbf n_a,\mathbf n_b,\mathbf n_{ab}\}
\]
in the collinear case.
For a soft particle, the measure changes by
\(\varepsilon\delta_{\mathbf m}\), so the change of the detector value is
\(\varepsilon f(\mathbf m)\), whose absolute value is bounded by
\(\varepsilon\|f\|_\infty\).  For a collinear recombination, the change is
\[
  E_a f(\mathbf n_a)+E_b f(\mathbf n_b)
  -(E_a+E_b)f(\mathbf n_{ab}).
\]
Writing this as
\[
  E_a\bigl(f(\mathbf n_a)-f(\mathbf n_{ab})\bigr)
  +
  E_b\bigl(f(\mathbf n_b)-f(\mathbf n_{ab})\bigr)
\]
and using the Lipschitz constant of \(f\) on the round sphere gives
\[
  |\Delta\widehat{\mathcal E}_X(f)|
  \leq
  E_a\operatorname{Lip}(f)d(\mathbf n_a,\mathbf n_{ab})
  +
  E_b\operatorname{Lip}(f)d(\mathbf n_b,\mathbf n_{ab}),
\]
which is bounded by the displayed diameter estimate.  Thus the smeared
detector is continuous for the weak calorimetric topology generated by
finite-energy measures tested against \(C^1(S^2)\) functions: soft additions
vanish with their energy, and collinear recombinations vanish uniformly on
bounded-energy sectors as the angular diameter tends to zero.

The same estimate applies to products
\(\prod_{\ell=1}^k\widehat{\mathcal E}_X(f_\ell)\) on any fixed
finite-energy sector by the elementary telescoping identity
\[
  \prod_{\ell=1}^k A_\ell-\prod_{\ell=1}^k B_\ell
  =
  \sum_{j=1}^k
  \left(\prod_{\ell<j}A_\ell\right)
  (A_j-B_j)
  \left(\prod_{\ell>j}B_\ell\right),
\]
with \(A_\ell=\widehat{\mathcal E}_{X'}(f_\ell)\) and
\(B_\ell=\widehat{\mathcal E}_X(f_\ell)\).  Thus smeared energy correlators
are continuous functions of the calorimetric measure for soft and collinear
degenerations.  Perturbative finiteness then has a precise theorem boundary:
at each fixed order in a regulator for which the Kinoshita--Lee--Nauenberg
cancellation and QCD factorization hold for the chosen inclusive channel,
measurement functions with the above continuity produce finite coefficients
after summing over degenerate final states and virtual corrections.  The
operator definition \eqref{eq:energy-correlator-definition} is the
nonperturbative observable; perturbation theory computes its short-distance
expansion in the regime where asymptotic freedom supplies a controlled small
coupling.

The unsmeared product
\(\widehat{\mathcal E}(\mathbf n_1)\cdots
\widehat{\mathcal E}(\mathbf n_k)\) is a distribution.  Coincident directions
carry contact singularities because several detector insertions can act on the
same hadron.  The separated-angle correlator is obtained by testing against
functions supported away from the coincidence diagonals in \((S^2)^k\); fully
smeared correlators include the contact terms determined by the detector
operator product itself.

\subsection{Energy-Energy Correlation in Electron-Positron Annihilation}

Let a color-singlet current create a state of total center-of-mass energy
\(Q\).  For each physical hadronic final state \(X\), write
\[
  z_r=\frac{p_r^0}{Q},
  \qquad
  \cos\chi_{rs}=
  \widehat{\mathbf p}_r\cdot\widehat{\mathbf p}_s .
\]
The normalized two-point energy-energy correlator is the distribution
\begin{equation}
  \mathrm{EEC}(\chi)
  =
  \frac{1}{\sigma_{\mathrm{tot}}}
  \sum_X
  \int \dd\sigma_X\,
  \sum_{r,s\in X}
  z_r z_s\,
  \delta(\cos\chi-\cos\chi_{rs}),
  \label{eq:qcd-eec-definition}
\end{equation}
where \(\dd\sigma_X\) is the differential cross-section measure for the
exclusive hadronic channel \(X\), and the sum is a direct integral over
continuous phase space.  The terms \(r=s\) are the coincident-detector contact
part at \(\chi=0\).  A convention that removes this contact part defines a
different distribution at \(\chi=0\) and the same separated-angle function for
\(0<\chi<\pi\).

Equation~\eqref{eq:qcd-eec-definition} is the detector form of the
Basham--Brown--Ellis--Love observable.  Its perturbative expansion in
asymptotically free QCD is a calculation of the current-created state and of
the detector insertions in the short-distance region.  The back-to-back limit
\(\chi\to\pi\) is controlled by soft and collinear radiation constrained by
transverse momentum conservation; the small-angle limit is controlled by a
collinear operator product of energy-flow insertions.  Both limits require
resummation because the separated-angle distribution contains large logarithms
of \(1+\cos\chi\) or \(1-\cos\chi\), respectively.

\paragraph{EEC moment sum rules and the contact term.}
Let \(X\) be a physical final state in the center-of-mass frame of a
color-singlet source, with total energy \(Q\) and total spatial momentum zero.
Set
\[
  z_r=\frac{p_r^0}{Q},
  \qquad
  \mathbf n_r=\frac{\mathbf p_r}{|\mathbf p_r|},
  \qquad
  \zeta_{rs}=\mathbf n_r\cdot\mathbf n_s .
\]
Define the eventwise two-point calorimetric distribution in
\(\zeta=\cos\chi\) by
\begin{equation}
  E_{2,X}(\zeta)
  =
  \sum_{r,s\in X}z_r z_s\,\delta(\zeta-\zeta_{rs}) .
  \label{eq:eventwise-eec-zeta-distribution}
\end{equation}
Then, as distributional moments on \([-1,1]\),
\begin{equation}
  \int_{-1}^{1} E_{2,X}(\zeta)\,\dd\zeta=1,
  \qquad
  \int_{-1}^{1}\zeta\,E_{2,X}(\zeta)\,\dd\zeta=0 .
  \label{eq:eventwise-eec-first-sum-rules}
\end{equation}
The coincident-detector contact part at \(\zeta=1\) has weight
\begin{equation}
  E_{2,X}^{\rm contact}(\zeta)
  =
  \left(\sum_{r\in X}z_r^2\right)\delta(\zeta-1).
  \label{eq:eventwise-eec-contact-weight}
\end{equation}
Consequently the cross-section averaged EEC of
\eqref{eq:qcd-eec-definition}, including the \(r=s\) terms, satisfies the same
zeroth and first moment sum rules after division by \(\sigma_{\rm tot}\).

The total-energy condition gives
\[
  \sum_{r\in X}z_r=1.
\]
Testing \eqref{eq:eventwise-eec-zeta-distribution} against the constant
function \(1\) gives
\[
  \sum_{r,s}z_rz_s
  =
  \left(\sum_r z_r\right)^2
  =
  1.
\]
Testing against \(\zeta\) gives
\[
  \sum_{r,s}z_rz_s\,\mathbf n_r\cdot\mathbf n_s
  =
  \left|\sum_r z_r\mathbf n_r\right|^2.
\]
Since the source is in its center-of-mass frame,
\[
  \sum_r p_r^0\mathbf n_r=\sum_r\mathbf p_r=0,
\]
and the first moment vanishes.  The contact contribution is exactly the
subsum with \(r=s\), for which \(\zeta_{rr}=1\).  Averaging the eventwise
identities with the positive cross-section measure and dividing by
\(\sigma_{\rm tot}\) preserves the two moment identities.

\begin{remark}[Use of the sum rules]
\label{rem:eec-sum-rules-use}
The eventwise sum rules above are statements about the nonperturbative
detector observable, not about partons.  A fixed-order
calculation, a resummed endpoint formula, a lattice-inspired approximation, or
a parton shower that claims to represent the normalized EEC must specify how
the \(\zeta=1\) contact term is treated and must satisfy the moment identities
after the same convention is applied.  If the contact term is removed, the
remaining separated-angle distribution no longer has unit integral by itself;
the missing weight is \(\langle\sum_r z_r^2\rangle\).
\end{remark}

\paragraph{Legendre moments and calorimetric multipoles.}
The contact-inclusive two-detector distribution has stronger eventwise
constraints than the first two moments alone.  For every \(\ell\ge0\) define
\[
  M_{\ell,X}:=
  \int_{-1}^{1}P_\ell(\zeta)E_{2,X}(\zeta)\,\dd\zeta
  =
  \sum_{r,s\in X}z_rz_s P_\ell(\mathbf n_r\!\cdot\!\mathbf n_s).
\]
With the usual orthonormal complex spherical harmonics on \(S^2\), the
addition theorem gives
\[
  P_\ell(\mathbf n\!\cdot\!\mathbf m)
  =
  \frac{4\pi}{2\ell+1}
  \sum_{a=-\ell}^{\ell}
  Y_{\ell a}(\mathbf n)\overline{Y_{\ell a}(\mathbf m)} .
\]
Hence
\begin{equation}
  M_{\ell,X}
  =
  \frac{4\pi}{2\ell+1}
  \sum_{a=-\ell}^{\ell}
  \left|
    \sum_{r\in X} z_rY_{\ell a}(\mathbf n_r)
  \right|^2
  \ge0 .
  \label{eq:eventwise-eec-legendre-positivity}
\end{equation}
This positivity is a property of the positive calorimetric measure
\(\sum_r z_r\delta_{\mathbf n_r}\).  It is not a property of the
separated-angle density after the coincident-detector atom has been removed.
Indeed, for \(\ell=1\),
\[
  M_{1,X}
  =
  \sum_{r,s}z_rz_s\,\mathbf n_r\!\cdot\!\mathbf n_s
  =
  \left|\sum_r z_r\mathbf n_r\right|^2,
\]
so the color-singlet center-of-mass condition sets the full moment to zero;
the separated part alone then has first Legendre moment
\(-\sum_r z_r^2\).  The contact atom is therefore forced by positivity and
momentum conservation, not by a cosmetic endpoint convention.

The next moment displays the quadrupole coordinate.  Put
\[
  Q_{ij}(X):=
  \sum_{r\in X}z_r
  \left(n_r^in_r^j-\frac13\delta_{ij}\right).
\]
Then
\[
  \operatorname{tr}Q(X)^2
  =
  \sum_{r,s}z_rz_s
  \left((\mathbf n_r\!\cdot\!\mathbf n_s)^2-\frac13\right),
  \qquad
  M_{2,X}
  =
  \frac32\,\operatorname{tr}Q(X)^2
  \ge0 .
\]
Cross-section averaging preserves
\eqref{eq:eventwise-eec-legendre-positivity}.  Thus any perturbative,
resummed, shower, or finite-resolution representation of the normalized EEC
must say which endpoint/contact convention realizes these multipole
positivity constraints.  A separated-angle approximation that fits the bulk
shape but gives a negative contact-inclusive \(M_{\ell}\) is not a positive
detector measure.

\paragraph{Three-detector moment and contact ledger.}
The same detector definition gives a genuinely higher-point distribution
before any partonic approximation is made.  For the same event \(X\), define
the ordered three-detector distribution on the three angular invariants
\(\zeta_{12},\zeta_{13},\zeta_{23}\) by
\begin{equation}
\begin{aligned}
  E_{3,X}(\zeta_{12},\zeta_{13},\zeta_{23})
  &:=
  \sum_{r,s,t\in X}
  z_rz_sz_t\,
  \delta(\zeta_{12}-\mathbf n_r\!\cdot\!\mathbf n_s)\,
  \delta(\zeta_{13}-\mathbf n_r\!\cdot\!\mathbf n_t)          \\
  &\hspace{6.3em}\times
  \delta(\zeta_{23}-\mathbf n_s\!\cdot\!\mathbf n_t).
\end{aligned}
  \label{eq:eventwise-three-detector-distribution}
\end{equation}
Its support is not the whole cube \([-1,1]^3\), but the Gram-positive
region of triples of unit directions, plus the diagonal strata.  As moments
of this distribution,
\begin{equation}
  \int E_{3,X}=1,
  \qquad
  \int \zeta_{12}E_{3,X}
  =
  \int \zeta_{13}E_{3,X}
  =
  \int \zeta_{23}E_{3,X}
  =
  0
  \label{eq:eventwise-three-detector-moments}
\end{equation}
in a color-singlet center-of-mass event.  The proof is the same detector
algebra as above: the constant test gives
\((\sum_r z_r)^3=1\), while, for example,
\[
  \sum_{r,s,t}z_rz_sz_t\,\mathbf n_r\cdot\mathbf n_s
  =
  \left|\sum_r z_r\mathbf n_r\right|^2
  \left(\sum_t z_t\right)
  =
  0 .
\]
The other two pair moments are identical after relabeling the ordered
detectors.

The diagonal contact strata are now richer than in the EEC.  The completely
coincident stratum is
\begin{equation}
  \left(\sum_{r\in X}z_r^3\right)
  \delta(\zeta_{12}-1)\delta(\zeta_{13}-1)\delta(\zeta_{23}-1).
  \label{eq:eventwise-three-detector-total-contact}
\end{equation}
There are also three pair-coincidence strata.  For instance, the stratum
where detectors \(1\) and \(2\) hit the same hadron but detector \(3\) hits a
different hadron is
\begin{equation}
  \sum_{\substack{r,t\in X\\ r\neq t}}
  z_r^2z_t\,
  \delta(\zeta_{12}-1)
  \delta(\zeta_{13}-\zeta_{rt})
  \delta(\zeta_{23}-\zeta_{rt}),
  \qquad
  \zeta_{rt}=\mathbf n_r\cdot\mathbf n_t,
  \label{eq:eventwise-three-detector-pair-contact}
\end{equation}
with two analogous terms for \(1=3\) and \(2=3\).  Removing all these
diagonal strata defines a separated three-detector observable.  It does not
by itself obey \eqref{eq:eventwise-three-detector-moments}; the missing
weights are precisely the contact coordinates above.  Thus higher-point
energy correlators inherit the same lesson as the EEC: endpoint/contact
coordinates are part of the detector observable, not optional decorations
added after a partonic calculation.

\subsection{Light-Ray Interpretation and Endpoint Regimes}

The detector density \(\widehat{\mathcal E}(\mathbf n)\) is the QCD analogue
of the stress-tensor light transform developed in
Chapter~\ref{ch:cft-light-ray-energy-correlators}.  In a conformal theory the
light transform is an operator-valued distribution controlled by Lorentzian
conformal representation theory.  In QCD the coupling runs, so the same
formula is not a conformal operator identity.  It remains an asymptotic
detector definition in the physical Hilbert space, and the short-distance
limits are computed by matching the detector product onto operators whose
renormalization group evolution is controlled by QCD anomalous dimensions.

\paragraph{Proof-status architecture for the EEC program.}
The exact detector identities above and the endpoint formulae below live at
different levels of the argument.  First, the physical detector layer starts
from the stress-tensor flux limit or, on the outgoing scattering
representation, from multiplication by the calorimetric measure.  This layer
fixes positivity, total-energy moments, diagonal contact strata, multipole
positivity, and the distinction between ensemble cumulants and eventwise
contacts.  Second, the conformal fixed-point layer is the light-ray
operator framework of Chapter~\ref{ch:cft-light-ray-energy-correlators}: it
supplies representation-theoretic homogeneity, finite light-ray charts,
diagonal-contact extensions, and endpoint pushforwards only under the
Lorentzian analyticity and growth hypotheses stated there.  Third, the QCD
small-angle layer replaces fixed-point light-ray data by renormalized
Wilson-line/light-ray operator coordinates with ultraviolet and rapidity
subtractions, protected moment rows, finite chart changes, and projected
curvature or omitted-channel remainders.  Fourth, the perturbative endpoint
layer computes coefficients, anomalous dimensions, contact counterterms, and
matching functions in one declared scheme; the tree-level collinear
coefficients and cusp-log flatness check are tests of this layer, not the
full all-order light-ray OPE/mixing theorem.  Finally, the observable
prediction layer glues the separated bulk, small-angle endpoint,
back-to-back endpoint, and explicit endpoint atoms so that the detector
moment identities remain true.

Thus the dependency chain is
\[
  \begin{aligned}
  \text{detector measure and contacts}
  &\Longrightarrow
  \text{CFT or QCD light-ray endpoint chart} \\
  &\Longrightarrow
  \text{renormalized mixing/factorization calculation} \\
  &\Longrightarrow
  \text{endpoint gluing and observable prediction}.
  \end{aligned}
\]
Skipping any arrow changes the mathematical claim.  A separated-angle
fixed-order distribution without contact data is not the normalized detector
observable; a finite retained light-ray block without a curvature or
omitted-channel estimate is not an endpoint theorem; and a resummed endpoint
formula without the gluing and moment ledger is not yet a prediction for the
full normalized EEC on \([-1,1]\).

For \(0<\chi<\pi\), the EEC is a separated-angle distribution.  Two endpoint
regions have different expansions.  In the small-angle region set
\[
  \theta^2:=2(1-\cos\chi)\ll1 .
\]
The product of two nearby detector insertions is matched onto a sum of
collinear light-ray operators:
\begin{equation}
  \widehat{\mathcal E}(\mathbf n_1)\widehat{\mathcal E}(\mathbf n_2)
  \sim
  \sum_A C_A(\theta,Q,\mu)\,
  \mathbb O_A(\mathbf n,\mu),
  \qquad
  \mathbf n_1,\mathbf n_2\to\mathbf n .
  \label{eq:qcd-eec-light-ray-ope-form}
\end{equation}
Here \(\mathbb O_A\) are light-ray operators in the collinear direction,
including twist-two energy-flow moments and their descendants in the
appropriate channel.  The symbol \(\sim\) denotes an asymptotic expansion
after smearing against angular test functions supported at
\(\theta\ll1\).  The coefficient functions \(C_A\) are perturbatively
computable at large transverse scale \(Q\theta\), and the evolution of
\(\mathbb O_A\) resums logarithms of \(\theta\).

In the back-to-back region set
\[
  z:=\frac{1-\cos\chi}{2},
  \qquad
  q_T^2:=Q^2(1-z)
  =
  \frac{Q^2}{2}(1+\cos\chi)\ll Q^2 .
\]
Momentum conservation forces the two measured energy deposits to recoil
against soft and collinear radiation of small total transverse momentum.  The
factorized form is the transverse-momentum resummation analogue
\begin{equation}
  \frac{\dd\sigma}{\dd q_T^2}
  =
  H(Q,\mu)
  \int \frac{\dd^2 b}{(2\pi)^2}\,
  \ee^{\ii b\cdot q_T}\,
  J_n(b,\mu,\nu)\,
  J_{\bar n}(b,\mu,\nu)\,
  S(b,\mu,\nu)
  +
  O(q_T/Q),
  \label{eq:qcd-eec-back-to-back-factorization}
\end{equation}
for a color-singlet source and in a rapidity-renormalization scheme with
rapidity scale \(\nu\).  The hard, jet, and soft functions have coupled
\(\mu\)- and \(\nu\)-evolution equations; solving them resums Sudakov
logarithms of \(q_T/Q\).  The power remainder and the absence or treatment of
Glauber exchanges are part of the factorization hypothesis.

Here and below \(\alpha_s(\mu):=g^2(\mu)/(4\pi)\) in the trace-delta gauge
coordinate of this chapter.  Whenever a coefficient is proportional to
\(\alpha_s C_F\), it is invariant under conversion to the common half-trace
generator convention because \(g_{\rm ht}^2C_F^{\rm ht}=g^2C_F\), as in
\eqref{eq:pnrqcd-tree-coulomb-potential}.

The leading double logarithm is fixed already by the lightlike cusp anomalous
dimension.  Let \(b=|b|\) in \eqref{eq:qcd-eec-back-to-back-factorization},
set
\[
  b_{\rm F}:=2\ee^{-\gamma_E},\qquad
  \mu_b:=\frac{b_{\rm F}}{b},\qquad
  L_b:=\log\frac{Q^2}{\mu_b^2}
  =
  \log\frac{Q^2b^2}{b_{\rm F}^2}.
\]
The constant \(b_{\rm F}\) is a Fourier-convention constant, not a new
physical scale.  In the trace-delta convention of this chapter the one-loop
quark cusp anomalous dimension is
\begin{equation}
  \Gamma_{\rm cusp}^q(g)
  =
  \frac{g^2C_F}{4\pi^2}+O(g^4)
  =
  \frac{\alpha_s C_F}{\pi}+O(\alpha_s^2),
  \label{eq:qcd-quark-cusp-one-loop-trace-delta}
\end{equation}
where \(C_F=(N_c^2-1)/N_c\).

Assume the back-to-back factorization datum
\eqref{eq:qcd-eec-back-to-back-factorization} for a color-singlet
quark-current event shape, with the hard scale \(Q\), the impact-parameter
scale \(\mu_b=b_{\rm F}/b\), and a subtraction scheme in which the leading
large logarithms are generated by the quark cusp anomalous dimension
\eqref{eq:qcd-quark-cusp-one-loop-trace-delta}.  If running of \(g\) and all
non-cusp anomalous dimensions are omitted, the leading-logarithmic
impact-parameter kernel has the form
\begin{equation}
  W_{\rm LL}(b,Q)
  =
  W(b,\mu_b)\,
  \exp\!\left[
    -\frac12\,\Gamma_{\rm cusp}^q(g)\,L_b^2
  \right].
  \label{eq:qcd-eec-leading-sudakov-factor}
\end{equation}
Equivalently, at one loop,
\[
  W_{\rm LL}(b,Q)
  =
  W(b,\mu_b)\,
  \exp\!\left[
    -\frac{\alpha_s C_F}{2\pi}L_b^2
    +O(\alpha_s^2L_b^2)
  \right].
\]
Its content is the fixed-coupling leading-logarithmic component of the
factorization formula.  The factorized expression is an impact-parameter
representation of the
recoil constraint.  The hard matching is naturally made at scale \(Q\), while
the transverse-distance functions are naturally evaluated at scale
\(\mu_b\).  Cusp evolution between these two scales gives the logarithmic
exponent
\[
  S_{\rm LL}(Q,b)
  =
  \int_{\mu_b^2}^{Q^2}
  \frac{\dd\kappa^2}{\kappa^2}\,
  \Gamma_{\rm cusp}^q(g)\,
  \log\frac{Q^2}{\kappa^2}
\]
when \(g\) is held fixed and non-cusp anomalous dimensions are discarded.
Changing variables to
\[
  u=\log\frac{Q^2}{\kappa^2}
\]
turns the integral into
\[
  S_{\rm LL}(Q,b)
  =
  \Gamma_{\rm cusp}^q(g)
  \int_0^{L_b}u\,\dd u
  =
  \frac12\,\Gamma_{\rm cusp}^q(g)L_b^2.
\]
The kernel in \eqref{eq:qcd-eec-back-to-back-factorization} is suppressed by
the Sudakov no-resolved-recoil factor between the two scales; this fixes the
minus sign in \(\exp[-S_{\rm LL}]\) in the chosen evolution convention.
Substituting
\eqref{eq:qcd-quark-cusp-one-loop-trace-delta} gives the displayed one-loop
form.

\begin{remark}[Scope of the Sudakov formula]
\label{rem:qcd-eec-sudakov-scope}
The fixed-coupling display above supplies only the cusp-controlled double
logarithm inside a full back-to-back EEC theorem.  A complete perturbative
prediction must specify the rapidity regulator, the hard, jet, and soft
functions, their finite scheme constants, matching to the fixed
separated-angle distribution, running-coupling improvement, and the leading
power corrections in \(q_T/Q\).  The calculation isolates the universal
cusp-controlled double logarithm that any such consistent factorization scheme
must reproduce.
\end{remark}

\paragraph{Tested recoil chart and the large-\(b\) boundary.}
The impact-parameter formula becomes the measured contribution
\(\mathcal P_\tau^{\rm rec}\) in
\eqref{eq:qcd-measured-eec-factorization-budget} only after it has been
turned into a detector-test functional.  One must first choose a
back-to-back endpoint window
\(0<q_T<q_0\) and pull angular tests back by
\begin{equation}
  \varphi_Q(q_T^2)
  :=
  \varphi\!\left(-1+\frac{2q_T^2}{Q^2}\right),
  \qquad
  q_T^2=\frac{Q^2}{2}(1+\zeta).
  \label{eq:qcd-eec-recoil-test-pullback}
\end{equation}
This is the pushforward of the standard EEC variable
\(z=(1-\zeta)/2\) together with \(1-z=q_T^2/Q^2\).  Equivalently,
\(\dd\sigma/\dd\zeta=(Q^2/2)\,\dd\sigma/\dd q_T^2\) at
\(q_T^2=Q^2(1+\zeta)/2\); in test-function form the same Jacobian is already
accounted for by pairing the \(q_T^2\)-distribution with
\(\varphi_Q(q_T^2)\).  The open recoil chart is then the distribution on
\(0<q_T<q_0\) obtained by pairing \(\varphi_Q\) with the Fourier kernel in
\eqref{eq:qcd-eec-back-to-back-factorization}, after the overlap with the
fixed-order bulk chart has been subtracted.  Endpoint support at \(q_T=0\), or
equivalently \(\zeta=-1\), is carried by the separate endpoint atom
\(K_-^\tau\) in
\eqref{eq:qcd-measured-eec-factorization-budget}.  This separation is
essential because a Born-level back-to-back contribution, a perturbative
recoil singularity, and a nonperturbative endpoint model are different
coordinates of the same measured distribution.

The Fourier integral also has a physical boundary absent from the displayed
leading logarithm.  Perturbation theory controls the factorized kernel only
while all transverse-distance functions are evaluated at scales safely above
\(\Lambda_{\rm QCD}\).  Large impact parameters \(b\gtrsim
\Lambda_{\rm QCD}^{-1}\), or bins with \(q_T\) of hadronic size, require a
nonperturbative shape function, a fitted large-\(b\) prescription, or an
excluded-bin remainder.  Changing from one such prescription to another
changes the long-distance input paired with the detector test, rather than
merely reshuffling the cusp logarithm.

\begin{controlledapproximation}[Tested back-to-back EEC recoil chart]
\label{ca:qcd-eec-tested-back-to-back-recoil}
A back-to-back contribution to a measured calorimetric or selected-track EEC
is controlled only after the following data have been specified:
\begin{enumerate}
  \item the endpoint window, angular test space, and pullback
  \eqref{eq:qcd-eec-recoil-test-pullback};
  \item the hard, jet, and soft functions in one ultraviolet and rapidity
  scheme, including the track-function or fragmentation lift when
  \(\tau\ne1\);
  \item the running-coupling, non-cusp, and rapidity evolution solving the
  same factorization equations as the fixed-order matching coefficients;
  \item the large-\(b\) prescription or shape input, together with the status
  of Glauber contributions and ordinary \(q_T/Q\) power corrections;
  \item the overlap subtraction with the separated-angle bulk chart and the
  endpoint atom convention for \(K_-^\tau\).
\end{enumerate}
If these choices give a recoil functional
\(\mathcal P_{\tau}^{\rm rec}[\varphi;Q]\), its contribution to the measured
residual in \eqref{eq:qcd-measured-eec-residual-bound} must be bounded in the
detector-test topology by a decomposition of the form
\begin{equation}
  \bigl|R_{\tau,Q}^{\rm rec}[\varphi]\bigr|
  \leq
  \epsilon_{\rm fac}^{\rm rec}(\varphi)
  +
  \epsilon_{\rm evo}^{\rm rec}(\varphi)
  +
  \epsilon_{\rm match}^{\rm rec}(\varphi)
  +
  \epsilon_{\rm large\text{-}b}^{\rm rec}(\varphi)
  +
  \epsilon_{\rm Glauber}^{\rm rec}(\varphi)
  +
  \epsilon_{\rm pow}^{\rm rec}(\varphi).
  \label{eq:qcd-eec-tested-recoil-residual-bound}
\end{equation}
Here \(\epsilon_{\rm fac}^{\rm rec}\) is the leading-power factorization
defect, \(\epsilon_{\rm evo}^{\rm rec}\) is the perturbative evolution and
truncation error, \(\epsilon_{\rm match}^{\rm rec}\) includes fixed-order
matching and overlap-subtraction errors, \(\epsilon_{\rm large\text{-}b}^{\rm rec}\)
is the nonperturbative impact-parameter contribution or excluded-bin error,
\(\epsilon_{\rm Glauber}^{\rm rec}\) records any failure or hypothesis of
Glauber cancellation, and \(\epsilon_{\rm pow}^{\rm rec}\) is the remaining
power correction.  The leading Sudakov factor
\eqref{eq:qcd-eec-leading-sudakov-factor} supplies only the cusp-controlled
piece of the evolution item.  The recoil contribution to a measured EEC bin
additionally requires the tested chart, endpoint matching, large-\(b\)
input, and residual data listed above.
\end{controlledapproximation}

\begin{lemma}[Tree-level collinear coefficient for the separated EEC]
\label{lem:qcd-eec-tree-collinear-coefficient}
Use the standard four-dimensional, azimuth-integrated, final-state collinear
splitting convention
\begin{equation}
  \dd P_{a\to bc}
  =
  \frac{\alpha_s}{2\pi}\,
  P_{a\to bc}(x)\,\dd x\,\frac{\dd\theta^2}{\theta^2},
  \qquad 0<x<1,
  \label{eq:final-state-collinear-splitting-convention}
\end{equation}
where \(x\) and \(1-x\) are the daughter energy fractions and \(\theta\) is
their small opening angle.  The real-emission contribution of this resolved
splitting to the ordered separated-angle EEC is weighted by
\begin{equation}
  w_{\rm EEC}(x)=2x(1-x).
  \label{eq:eec-ordered-collinear-weight}
\end{equation}
Consequently, for \(\zeta=\cos\chi\) and
\(\theta^2=2(1-\zeta)+O((1-\zeta)^2)\), the leading-power separated-angle
collinear density generated by a parent of type \(a\) has the form
\begin{equation}
  \frac{\dd P_a^{\rm EEC}}{\dd\zeta}
  =
  \frac{\alpha_s}{2\pi}\,
  \frac{\Gamma_a^{\rm EEC}}{1-\zeta}
  +
  \text{terms less singular as }\zeta\to1,
  \label{eq:eec-tree-collinear-density}
\end{equation}
with resolved real-kernel coefficients
\begin{equation}
\begin{aligned}
  \Gamma_q^{\rm EEC}
  &=
  \int_0^1 2x(1-x)\,
  C_F\frac{1+x^2}{1-x}\,\dd x
  =
  \frac32\,C_F,                                      \\
  \Gamma_{g\to gg}^{\rm EEC}
  &=
  \int_0^1 2x(1-x)\,
  2C_A\left[
    \frac{x}{1-x}+\frac{1-x}{x}+x(1-x)
  \right]\dd x
  =
  \frac{14}{5}\,C_A,                                 \\
  \Gamma_{g\to q\bar q}^{\rm EEC}
  &=
  \int_0^1 2x(1-x)\,
  T_F\bigl(x^2+(1-x)^2\bigr)\dd x
  =
  \frac15\,T_F
\end{aligned}
\label{eq:eec-real-collinear-coefficients}
\end{equation}
per quark flavor in the last line.  The endpoint singularities of the real
splitting kernels are integrable in these coefficients because the detector
weight \(2x(1-x)\) vanishes when either daughter carries zero energy.
\end{lemma}

\begin{proof}
For one parent of energy \(E\) splitting into two collinear daughters of
energies \(xE\) and \((1-x)E\), the ordered two-detector contribution at the
nonzero opening angle consists of the two ordered pairs \((b,c)\) and
\((c,b)\).  Its energy weight divided by \(E^2\) is therefore
\[
  x(1-x)+(1-x)x=2x(1-x),
\]
which proves \eqref{eq:eec-ordered-collinear-weight}.  The coincident
daughter terms have contact weight
\[
  x^2+(1-x)^2,
\]
and the identity
\[
  x^2+(1-x)^2+2x(1-x)=1
\]
is the local version of the zeroth-moment sum rule in
the eventwise EEC moment identities above: a collinear splitting redistributes
the parent's detector weight between the contact point and the separated small
angle without changing the total two-detector weight.

At leading collinear power the probability measure for the resolved splitting
is \eqref{eq:final-state-collinear-splitting-convention}.  Since
\[
  1-\zeta=1-\cos\theta=\frac{\theta^2}{2}+O(\theta^4),
\]
the positive measure \(\dd\theta^2/\theta^2\), pulled back to the
\(\zeta\)-line near \(\zeta=1\), is
\[
  \frac{\dd\theta^2}{\theta^2}
  =
  \frac{\dd\zeta}{1-\zeta}+O(1)\,\dd\zeta ,
\]
where the equality is an equality of positive densities on intervals
\(\zeta<1\), not an oriented one-form identity.
Multiplying the splitting probability by the ordered EEC weight and retaining
only the leading distribution as \(\zeta\to1\) gives
\eqref{eq:eec-tree-collinear-density} with
\[
  \Gamma_a^{\rm EEC}
  =
  \sum_{a\to bc}
  \int_0^1 2x(1-x)P_{a\to bc}(x)\,\dd x .
\]
Substituting the real tree-level kernels displayed in
\eqref{eq:eec-real-collinear-coefficients} reduces the three coefficients to
elementary polynomial integrals:
\[
  2C_F\int_0^1 x(1+x^2)\,\dd x=\frac32C_F,
\]
\[
  4C_A\int_0^1
  \bigl[x^2+(1-x)^2+x^2(1-x)^2\bigr]\dd x
  =
  \frac{14}{5}C_A,
\]
and
\[
  2T_F\int_0^1
  x(1-x)\bigl(x^2+(1-x)^2\bigr)\dd x
  =
  \frac15T_F .
\]
This proves \eqref{eq:eec-real-collinear-coefficients}.
\end{proof}

\begin{remark}[What the tree coefficient does and does not prove]
\label{rem:eec-tree-collinear-coefficient-status}
Lemma~\ref{lem:qcd-eec-tree-collinear-coefficient} is a derivation of
the first resolved real-emission coefficient in the small-angle expansion
under the stated collinear splitting convention.  It is not the full
small-angle EEC theorem.  Beyond this leading coefficient, one must define the
renormalized light-ray operator basis, include virtual terms and contact
conventions, solve the mixing problem, and match the endpoint expansion to
the separated-angle distribution in a fixed regulator and subtraction scheme.
The exact observable remains the detector distribution
\eqref{eq:qcd-eec-definition}.
\end{remark}

\paragraph{Small-angle light-ray OPE structure.}
The symbolic matching formula \eqref{eq:qcd-eec-light-ray-ope-form} becomes
a mathematical statement only after the operator coordinates, endpoint
variable, regulator, and contact convention have been declared.  Let
\[
  \rho:=1-\zeta=1-\cos\chi,\qquad
  \rho\simeq \frac{\theta^2}{2}
\]
in the small-angle region.  A renormalized QCD light-ray OPE structure for this
endpoint consists of the following ingredients:
\[
  \mathcal D_{\rm EEC}^{\rm coll}
  =
  \bigl(
    \mathfrak L_{\mathbf n},\{\mathbb O_A\},
    \mathcal S_{\rm UV},\mathcal S_{\rm rap},
    \gamma_{AB},\eta_{AB},C_A
  \bigr).
\]
Here \(\mathfrak L_{\mathbf n}\) is a linear space of gauge-invariant
renormalized light-ray operator-valued distributions in the null direction
specified by \(\mathbf n\).  The indexed family
\(\{\mathbb O_A(\mathbf n;\mu,\nu)\}\) is a chosen coordinate basis of this
space, including the leading twist energy-flow moments and any operators
that mix with them under the declared symmetries.  The symbols
\(\mathcal S_{\rm UV}\) and \(\mathcal S_{\rm rap}\) denote, respectively,
the ultraviolet subtraction scheme and, when a lightlike Wilson-line or
rapidity singularity is present, the rapidity subtraction scheme.  The scale
\(\mu\) is the ultraviolet renormalization scale and \(\nu\) is the rapidity
scale; if a scheme has no separate rapidity regulator, all \(\nu\)-dependent
objects below are omitted.  The kernels \(\gamma_{AB}\) and \(\eta_{AB}\)
are distributional mixing kernels in whatever momentum-fraction, Mellin, or
angular variables label the basis.  The coefficient \(C_A(\rho,Q,\mu,\nu)\)
acts on smooth endpoint test functions in \(\rho\), with its separated-angle
density, endpoint subtractions, and \(\delta(\rho)\) contact coefficient all
specified by the same subtraction scheme.

With column-vector operator convention, the renormalization equations are
\begin{equation}
  \mu\frac{\dd}{\dd\mu}\mathbb O_A
  =
  -\sum_B \gamma_{AB}\otimes\mathbb O_B,\qquad
  \nu\frac{\dd}{\dd\nu}\mathbb O_A
  =
  -\sum_B \eta_{AB}\otimes\mathbb O_B ,
  \label{eq:qcd-eec-light-ray-operator-rg}
\end{equation}
where \(\otimes\) denotes the convolution product in the operator labels and
matrix multiplication in the discrete labels.  Since the product of
detectors is a physical operator-valued distribution after the detector
observable is normalized, the endpoint expansion
\[
  \widehat{\mathcal E}(\mathbf n_1)\widehat{\mathcal E}(\mathbf n_2)
  \simeq
  \sum_A C_A\otimes \mathbb O_A
\]
must be independent of \(\mu\) and of \(\nu\).  Differentiating this
identity and using \eqref{eq:qcd-eec-light-ray-operator-rg} gives the
coefficient equations
\begin{equation}
  \mu\frac{\dd}{\dd\mu}C_A
  =
  \sum_B C_B\otimes\gamma_{BA},\qquad
  \nu\frac{\dd}{\dd\nu}C_A
  =
  \sum_B C_B\otimes\eta_{BA}.
  \label{eq:qcd-eec-light-ray-coefficient-rg}
\end{equation}
This sign and transpose convention fixes the cancellation between the
anomalous dimension of the coefficient row and the anomalous dimension of the
operator column.  In a Mellin basis where the convolution diagonalizes,
\eqref{eq:qcd-eec-light-ray-coefficient-rg} reduces to ordinary matrix
multiplication in the remaining operator indices.

The separated-angle real-emission coefficient in
Lemma~\ref{lem:qcd-eec-tree-collinear-coefficient} supplies the positive
density on \(\rho>0\).  The renormalized endpoint coefficient also contains
endpoint subtractions and coincident-detector contact data.  To state this
coefficient, fix a small but finite endpoint interval
\(0<\rho<\rho_0\).  The plus distribution on this interval is defined by
\begin{equation}
  \int_0^{\rho_0}
  \left[\frac1\rho\right]_{\rho_0,+}\varphi(\rho)\,\dd\rho
  :=
  \int_0^{\rho_0}
  \frac{\varphi(\rho)-\varphi(0)}{\rho}\,\dd\rho
  \label{eq:qcd-eec-rho-plus-definition}
\end{equation}
for every smooth test function \(\varphi\) on \([0,\rho_0]\).  Therefore the
first singular real coefficient
\[
  \frac{\alpha_s}{2\pi}\,
  \frac{\Gamma_a^{\rm EEC}}{\rho}
\]
is promoted, in a renormalized endpoint coefficient, to the distributional
form
\begin{equation}
  C_a^{(1)}(\rho)
  =
  \frac{\alpha_s}{2\pi}\,
  \Gamma_a^{\rm EEC}
  \left[\frac1\rho\right]_{\rho_0,+}
  +
  K_a^{(1)}(\rho_0,\mu,\nu)\,\delta(\rho)
  +
  C_{a,{\rm reg}}^{(1)}(\rho).
  \label{eq:qcd-eec-renormalized-endpoint-coefficient}
\end{equation}
The constant \(K_a^{(1)}\) and the regular distribution
\(C_{a,{\rm reg}}^{(1)}\) are scheme- and matching-dependent.  The plus
prescription carries the coincident-detector subtraction that makes the
endpoint distribution act on test functions with finite endpoint limits.  The
remaining finite \(\delta(\rho)\) coefficient is fixed after the virtual
terms, operator renormalization, and matching convention are specified in the
same scheme.  Thus the real coefficient \(\Gamma_a^{\rm EEC}\) is a
necessary local input to the small-angle OPE; the anomalous dimensions and
the full coefficient of a renormalized light-ray operator are determined by
the complete scheme-dependent endpoint datum.

The dependence on the endpoint resolution boundary is also distributional,
not verbal.  For \(0<\rho_a<\rho_b\), define
\[
  D_{\rho_*}(\rho)
  :=
  \left[\frac{\Theta(\rho_*-\rho)}{\rho}\right]_{+}
\]
by
\[
  \langle D_{\rho_*},\varphi\rangle
  =
  \int_0^{\rho_*}
  \frac{\varphi(\rho)-\varphi(0)}{\rho}\,\dd\rho
\]
for smooth endpoint tests.  Then, as distributions on \([0,\rho_b]\),
\begin{equation}
  D_{\rho_b}
  =
  D_{\rho_a}
  +
  \frac{\mathbf 1_{\rho_a<\rho<\rho_b}}{\rho}
  -
  \log\frac{\rho_b}{\rho_a}\,\delta(\rho).
  \label{eq:qcd-eec-plus-resolution-shift}
\end{equation}
Testing the right-hand side against \(\varphi\) gives
\[
  \int_0^{\rho_a}\frac{\varphi-\varphi(0)}{\rho}\,\dd\rho
  +
  \int_{\rho_a}^{\rho_b}\frac{\varphi(\rho)}{\rho}\,\dd\rho
  -
  \log\frac{\rho_b}{\rho_a}\varphi(0)
  =
  \int_0^{\rho_b}\frac{\varphi-\varphi(0)}{\rho}\,\dd\rho .
\]
Thus enlarging the endpoint chart by the annulus
\(\rho_a<\rho<\rho_b\) adds the ordinary separated-angle integral over that
annulus and simultaneously shifts the coincident-detector contact coordinate
by the logarithmic amount required to keep constant tests unchanged.  This is
the finite-resolution form of the endpoint contact convention: a statement
such as ``drop the contact term'' is incomplete until the endpoint resolution
boundary and the compensating \(\delta(\rho)\) coordinate have both been
declared.

The eventwise moment identity
\eqref{eq:eventwise-eec-first-sum-rules} imposes a useful check on this
bookkeeping.  Let \(M_A\) be the energy-sum functional on the chosen
light-ray operator basis, so that \(M_A\mathbb O_A\) extracts the component
that contributes to the zeroth EEC moment.  Conservation of total energy says
that this functional has zero anomalous dimension:
\begin{equation}
  \sum_A M_A\,\gamma_{AB}=0,
  \qquad
  \sum_A M_A\,\eta_{AB}=0
  \label{eq:qcd-eec-moment-functional-rg-invariance}
\end{equation}
in any scheme that preserves the detector normalization.  Combining
\eqref{eq:qcd-eec-light-ray-coefficient-rg} with
\eqref{eq:qcd-eec-moment-functional-rg-invariance} shows that the RG
evolution cannot change the zeroth moment; only the declared redistribution
between separated-angle weight and the \(\delta(\rho)\) contact term can
change the displayed representative.  This is why every perturbative EEC
formula must specify whether it includes the coincident-detector contact
term.  Without that convention, even a formally correct endpoint singularity
does not define the same distribution as the nonperturbative observable
\eqref{eq:qcd-eec-definition}.

\begin{proposition}[Timelike one-loop momentum sum rule for the EEC endpoint]
\label{prop:qcd-eec-timelike-momentum-sum-rule}
The protected row in
\eqref{eq:qcd-eec-moment-functional-rg-invariance} is visible already in the
one-loop final-state quark--gluon splitting kernels.  Let
\[
  D_0(x)=\left(\frac1{1-x}\right)_+
\]
with
\[
  \int_0^1D_0(x)\varphi(x)\,\dd x
  =
  \int_0^1\frac{\varphi(x)-\varphi(1)}{1-x}\,\dd x .
\]
In the trace-delta convention of this chapter, the leading timelike singlet
kernel for a daughter parton \(i\in\{q,g\}\) carrying energy fraction \(x\)
inside a parent \(j\in\{q,g\}\) may be written
\begin{align}
  P^{T,(0)}_{q\leftarrow q}(x)
  &=
  C_F\bigl(2D_0(x)-1-x+\tfrac32\delta(1-x)\bigr),
  \notag\\
  P^{T,(0)}_{g\leftarrow q}(x)
  &=
  C_F\,\frac{1+(1-x)^2}{x},
  \notag\\
  P^{T,(0)}_{q\leftarrow g}(x)
  &=
  2N_fT_F\bigl(x^2+(1-x)^2\bigr),
  \notag\\
  P^{T,(0)}_{g\leftarrow g}(x)
  &=
  2C_A\left[
    xD_0(x)+\frac{1-x}{x}+x(1-x)
  \right]
  +
  \frac{11C_A-4T_FN_f}{6}\delta(1-x).
  \label{eq:qcd-eec-timelike-one-loop-kernels}
\end{align}
Here \(q\) denotes the flavor-singlet quark-plus-antiquark energy-flow
coordinate; the factor \(2N_f\) in \(P^{T,(0)}_{q\leftarrow g}\) sums over
quarks and antiquarks.  Define the first energy-fraction moments
\[
  K_{i\leftarrow j}:=\int_0^1x\,P^{T,(0)}_{i\leftarrow j}(x)\,\dd x .
\]
Then
\begin{equation}
  K
  =
  \begin{pmatrix}
  -\frac43C_F & \frac23N_fT_F\\[4pt]
   \frac43C_F & -\frac23N_fT_F
  \end{pmatrix},
  \qquad
  (1\ \ 1)K=(0\ \ 0).
  \label{eq:qcd-eec-timelike-momentum-matrix}
\end{equation}
Thus the one-loop timelike mixing kernel redistributes energy between the
quark-singlet and gluon light-ray coordinates without changing the total
energy-flow row.  This is the perturbative kernel-level version of
\eqref{eq:qcd-eec-moment-functional-rg-invariance}.
\end{proposition}

\begin{proof}
The plus prescription gives
\[
  \int_0^1x^mD_0(x)\,\dd x
  =
  \int_0^1\frac{x^m-1}{1-x}\,\dd x
  =
  -H_m,
  \qquad H_0=0,\quad H_m=\sum_{r=1}^m\frac1r .
\]
For a quark parent,
\[
  \int_0^1xP^{T,(0)}_{q\leftarrow q}(x)\,\dd x
  =
  C_F\left(-2-\frac12-\frac13+\frac32\right)
  =
  -\frac43C_F,
\]
while
\[
  \int_0^1xP^{T,(0)}_{g\leftarrow q}(x)\,\dd x
  =
  C_F\int_0^1\bigl(1+(1-x)^2\bigr)\,\dd x
  =
  \frac43C_F .
\]
For a gluon parent,
\[
  \int_0^1xP^{T,(0)}_{q\leftarrow g}(x)\,\dd x
  =
  2N_fT_F
  \int_0^1x\bigl(x^2+(1-x)^2\bigr)\,\dd x
  =
  \frac23N_fT_F,
\]
and
\[
\begin{aligned}
  \int_0^1xP^{T,(0)}_{g\leftarrow g}(x)\,\dd x
  &=
  2C_A\left(-H_2+\frac12+\frac1{12}\right)
  +
  \frac{11C_A-4T_FN_f}{6} \\
  &=
  -\frac23N_fT_F .
\end{aligned}
\]
The two column sums are therefore zero.  The \(\delta(1-x)\) coefficient in
the gluon kernel is essential: omitting it would leave an uncancelled
\(-11C_A/6\) in the gluon-parent energy column.  Likewise, plus-prescribing
the quark pole is what turns the soft real-emission endpoint into the
negative quark-parent moment \(-4C_F/3\) needed to cancel the resolved gluon
daughter.  Both endpoint subtractions are therefore part of the detector
normalization mechanism, not optional notation.
\end{proof}

\paragraph{Finite renormalized light-ray mixing chart.}
The preceding equations describe the continuum endpoint expansion, whose full
operator basis is infinite and distributional.  In an actual calculation one
uses a finite coordinate chart on this space.  The following finite chart is
the algebraic core that every higher-order small-angle EEC calculation must
instantiate before its anomalous dimensions and matching coefficients can be
compared across schemes.

Fix a finite set of light-ray operator labels \(\mathcal I_L\), for example a
spin, twist, flavor, and derivative truncation, and let
\(\mathbb O^{\rm reg}_I(\Lambda,\nu_0)\), \(I\in\mathcal I_L\), be regulated
gauge-invariant light-ray composites with the endpoint and rapidity regulator
specified.  A finite renormalized chart consists of an invertible matrix
\(Z_{IA}(\Lambda,\mu,\nu;\mathcal R)\) and renormalized operator coordinates
defined by
\begin{equation}
  \mathbb O_A(\mu,\nu;\mathcal R)
  =
  \sum_{I\in\mathcal I_L}
  (Z^{-1})_{AI}(\Lambda,\mu,\nu;\mathcal R)\,
  \mathbb O_I^{\rm reg}(\Lambda,\nu_0),
  \label{eq:qcd-eec-finite-chart-renormalized-operators}
\end{equation}
after the regulator limit is taken in matrix elements in the declared test
space.  If the regulated coordinates are independent of \(\mu\) after the
regulator and bare parameters are held fixed, then differentiating
\eqref{eq:qcd-eec-finite-chart-renormalized-operators} gives
\begin{equation}
  \gamma
  =
  Z^{-1}\mu\frac{\dd Z}{\dd\mu},
  \qquad
  \mu\frac{\dd}{\dd\mu}\mathbb O=-\gamma\,\mathbb O ,
  \label{eq:qcd-eec-finite-chart-gamma-from-z}
\end{equation}
with column-vector notation.  A regulated coefficient row \(C^{\rm reg}\)
that pairs with the regulated operators as \(C^{\rm reg}\mathbb O^{\rm reg}\)
therefore gives the renormalized coefficient row
\begin{equation}
  C(\rho,Q,\mu,\nu;\mathcal R)
  =
  C^{\rm reg}(\rho,Q,\Lambda,\nu_0)\,
  Z(\Lambda,\mu,\nu;\mathcal R),
  \label{eq:qcd-eec-finite-chart-coefficient-renormalization}
\end{equation}
and hence
\begin{equation}
  \mu\frac{\dd C}{\dd\mu}=C\,\gamma .
  \label{eq:qcd-eec-finite-chart-coefficient-rg}
\end{equation}
This is the finite-matrix derivation of the sign and transpose convention in
\eqref{eq:qcd-eec-light-ray-operator-rg}--\eqref{eq:qcd-eec-light-ray-coefficient-rg}.
No physical statement is attached to \(Z\) or \(C\) separately: only the
paired distribution \(C\mathbb O\), together with its endpoint contact
coordinate, represents the detector product.

Finite scheme changes are coordinate changes on the same renormalized
light-ray space.  Let \(R(\mu,\nu)\) be an invertible finite matrix and set
\[
  \mathbb O'=R^{-1}\mathbb O,\qquad C'=C R .
\]
Then \(C'\mathbb O'=C\mathbb O\), while the anomalous-dimension matrix
becomes
\begin{equation}
  \gamma'
  =
  R^{-1}\gamma R
  +
  R^{-1}\mu\frac{\dd R}{\dd\mu},
  \label{eq:qcd-eec-finite-chart-scheme-gamma}
\end{equation}
and \(C'\) obeys \(\mu\dd C'/\dd\mu=C'\gamma'\).  Thus a quoted
``light-ray anomalous dimension'' is incomplete until its operator coordinate
chart is specified.  If \(M\mathbb O\) is the protected total-energy moment,
then \(M\gamma=0\).  In the changed coordinates \(M'=M R\) and
\[
  \mu\frac{\dd M'}{\dd\mu}=M'\gamma' .
\]
The stronger condition that the displayed row \(M\) is unchanged is therefore
not automatic; it is a restriction on admissible schemes, \(M R=M\).  This is
the finite-coordinate version of the normalization convention behind
\eqref{eq:qcd-eec-moment-functional-rg-invariance}.

\paragraph{Finite transport and two-scale consistency.}
The finite chart also makes precise what is meant by solving the light-ray
mixing problem.  Write
\[
  t=\log\mu,\qquad r=\log\nu .
\]
For a single ultraviolet scale, let \(U(t,t_0)\) be the matrix solution of
\begin{equation}
  \partial_tU(t,t_0)=U(t,t_0)\gamma(t),
  \qquad
  U(t_0,t_0)=\mathbf 1 .
  \label{eq:qcd-eec-light-ray-transport-u}
\end{equation}
When \(\gamma(t)\) at different scales does not commute,
\eqref{eq:qcd-eec-light-ray-transport-u} is the definition of the
path-ordered product; no choice of ordered exponential is meaningful without
the side on which \(\gamma\) acts.  With the row/column convention of
\eqref{eq:qcd-eec-finite-chart-gamma-from-z}--\eqref{eq:qcd-eec-finite-chart-coefficient-rg},
the transported coordinates are
\begin{equation}
  C(t)=C(t_0)U(t,t_0),
  \qquad
  \mathbb O(t)=U(t,t_0)^{-1}\mathbb O(t_0),
  \label{eq:qcd-eec-light-ray-transport-pair}
\end{equation}
and the paired endpoint distribution is independent of \(t\):
\[
  C(t)\mathbb O(t)=C(t_0)\mathbb O(t_0).
\]
Indeed, differentiating \(U^{-1}U=\mathbf 1\) gives
\[
  \partial_tU^{-1}=-\gamma(t)U^{-1},
\]
because \(\partial_tU=U\gamma\).  Thus
\[
  \partial_t\mathbb O(t)=-\gamma(t)\mathbb O(t),
  \qquad
  \partial_tC(t)=C(t)\gamma(t),
\]
and the two terms in \(\partial_t(C\mathbb O)\) cancel.  If \(M(t)\) is the
coordinate row representing the protected total-energy moment, its transport
law in a moving chart is
\begin{equation}
  \partial_tM(t)=M(t)\gamma(t).
  \label{eq:qcd-eec-light-ray-moment-transport}
\end{equation}
The convention \(M(t)=M(t_0)\) for all \(t\) is therefore equivalent to
\(M\gamma(t)=0\) in that fixed chart.  This is the all-order version of the
finite left-null-vector check in
\eqref{eq:qcd-eec-moment-functional-rg-invariance}.

If both ultraviolet and rapidity scales are present, the operator column is
declared to obey
\[
  \partial_t\mathbb O=-\gamma(t,r)\mathbb O,
  \qquad
  \partial_r\mathbb O=-\eta(t,r)\mathbb O .
\]
The two scale evolutions define a path-independent endpoint coordinate only
if the connection is flat on the chart:
\begin{equation}
  \partial_t\eta-\partial_r\gamma+[\gamma,\eta]=0 .
  \label{eq:qcd-eec-light-ray-flatness}
\end{equation}
This equation is not an optional aesthetic condition.  It is the equality of
mixed derivatives of the same renormalized operator coordinate.  To see the
sign, compute
\[
  \partial_r\partial_t\mathbb O
  =
  -(\partial_r\gamma)\mathbb O+\gamma\eta\,\mathbb O,
  \qquad
  \partial_t\partial_r\mathbb O
  =
  -(\partial_t\eta)\mathbb O+\eta\gamma\,\mathbb O .
\]
Equating the two expressions gives
\eqref{eq:qcd-eec-light-ray-flatness}.  In a finite chart with nonzero
curvature
\[
  \mathcal F_{tr}
  :=
  \partial_t\eta-\partial_r\gamma+[\gamma,\eta],
\]
transporting \(C\) and \(\mathbb O\) along two different paths in
\((\mu,\nu)\)-space produces different coordinate representatives unless the
curvature action is compensated by additional operators or by a prescribed
scheme-change wall.  A calculation that quotes separate \(\mu\)- and
\(\nu\)-anomalous dimensions for the EEC endpoint has therefore supplied a
consistent renormalized light-ray chart only after
\eqref{eq:qcd-eec-light-ray-flatness}, the protected moment row, and the
endpoint contact convention have been checked in the same scheme.

This paragraph gives the structural consistency condition.  It is not itself
a small-angle EEC loop calculation.  In a concrete rapidity-factorized scheme
one must display the regulated endpoint operator basis, compute the two
matrices \(\gamma(t,r)\) and \(\eta(t,r)\) with the same subtraction and
contact-term convention, and then verify
\(\partial_t\eta-\partial_r\gamma+[\gamma,\eta]=0\) on that basis.  If the
finite basis is a truncation of an infinite operator space, the check must
also state the projected curvature
\(P_L(\partial_t\eta-\partial_r\gamma+[\gamma,\eta])P_L\), the curvature
components that leave the retained subspace, and the contact-coordinate
terms that compensate endpoint scheme changes.  Vanishing only after
discarding an omitted operator is not flatness of the retained chart; it is
an approximation statement requiring an omitted-operator remainder.  The
finite matrix check recorded in the companion script verifies the sign and
row/column convention of the flatness equation, including the derivative
terms, but it does not replace this scheme-specific QCD calculation.

\paragraph{Projected curvature and omitted light-ray coordinates.}
The most common failure mode in a finite light-ray calculation is not an
incorrect commutator sign, but an omitted operator channel.  Let \(P_L\) be the
projection onto the retained light-ray labels and \(Q_L=1-P_L\).  In a fixed
chart write
\[
  \gamma_{LL}=P_L\gamma P_L,\qquad
  \gamma_{LQ}=P_L\gamma Q_L,\qquad
  \gamma_{QL}=Q_L\gamma P_L,
\]
and similarly for \(\eta\).  If the full connection is flat on the enlarged
operator space and \(P_L\) is scale-independent, then the curvature computed
from the retained block alone is
\begin{equation}
  \partial_t\eta_{LL}-\partial_r\gamma_{LL}
  +[\gamma_{LL},\eta_{LL}]
  =
  \eta_{LQ}\gamma_{QL}-\gamma_{LQ}\eta_{QL}.
  \label{eq:qcd-eec-projected-curvature-omitted-block}
\end{equation}
Indeed, insert \(P_L+Q_L=1\) between the two factors in the commutator part of
\eqref{eq:qcd-eec-light-ray-flatness}.  Thus a nonzero retained curvature is
not automatically an inconsistency of the full endpoint construction; it is an
explicit omitted-channel term.  A finite small-angle EEC calculation that keeps
only \(P_L\) must therefore either include the right-hand side of
\eqref{eq:qcd-eec-projected-curvature-omitted-block} in its remainder estimate
or enlarge the light-ray/contact-coordinate chart until the flatness check is
performed on the enlarged block.

This statement is coordinate covariant.  For a finite change of retained
coordinates \(R_L(t,r)\),
\[
  \mathbb O'_L=R_L^{-1}\mathbb O_L,\qquad C'_L=C_LR_L,
\]
one has
\[
  \gamma'_{LL}=R_L^{-1}\gamma_{LL}R_L+R_L^{-1}\partial_tR_L,
  \qquad
  \eta'_{LL}=R_L^{-1}\eta_{LL}R_L+R_L^{-1}\partial_rR_L,
\]
and the retained curvature transforms by conjugation,
\[
  \mathcal F'_{LL}=R_L^{-1}\mathcal F_{LL}R_L .
\]
Consequently a retained curvature cannot be removed by changing finite
coordinates inside the same retained space.  What can change is the split
between an operator block and endpoint contact coordinates, in which case the
flatness equation must be checked in the enlarged chart and the induced change
of contact distribution must be recorded in the endpoint convention.

\paragraph{Endpoint observable transport budget.}
The last step in a finite endpoint calculation is not the display of
\(\gamma\), \(\eta\), or a coefficient row.  It is the statement that the
retained chart gives a stable distribution when paired with endpoint detector
tests.  The following finite proposition is the algebraic budget behind that
statement.

\begin{proposition}[Finite endpoint observable transport budget]
\label{prop:qcd-eec-finite-endpoint-observable-transport-budget}
Fix a finite retained light-ray label set \(L\), an endpoint test
\(\varphi\), and one logarithmic scale \(t=\log\mu\).  Suppose the retained
small-angle endpoint contribution in a fixed state is represented by
\begin{equation}
  V_L(\varphi;t)
  =
  C_L(\varphi;t)\,O_L(t)
  +
  K_L(\varphi;t),
  \label{eq:qcd-eec-finite-endpoint-observable-chart}
\end{equation}
where \(C_L(\varphi;t)\) is a row of coefficient distributions evaluated on
\(\varphi\), \(O_L(t)\) is the column of retained light-ray matrix elements,
and \(K_L(\varphi;t)\) is the explicit endpoint contact coordinate.  Let
\(\gamma_{LL}(t)\) be the retained anomalous-dimension matrix, and let
\(D_L(\varphi;t)\) be the row produced by a scale-dependent change of
endpoint extension convention, such as moving an annulus into a plus-chart.
Assume
\begin{align}
  \partial_tO_L
  &=
  -\gamma_{LL}O_L+E_O,
  \label{eq:qcd-eec-endpoint-budget-operator}\\
  \partial_tC_L(\varphi)
  &=
  C_L(\varphi)\gamma_{LL}
  +D_L(\varphi)
  +E_C(\varphi),
  \label{eq:qcd-eec-endpoint-budget-coefficient}\\
  \partial_tK_L(\varphi)
  &=
  -D_L(\varphi)O_L+E_K(\varphi).
  \label{eq:qcd-eec-endpoint-budget-contact}
\end{align}
Then
\begin{equation}
  \partial_tV_L(\varphi;t)
  =
  E_C(\varphi;t)O_L(t)
  +
  C_L(\varphi;t)E_O(t)
  +
  E_K(\varphi;t).
  \label{eq:qcd-eec-endpoint-budget-residual}
\end{equation}
Consequently, for any interval \([t_0,t_1]\),
\begin{equation}
  \bigl|V_L(\varphi;t_1)-V_L(\varphi;t_0)\bigr|
  \leq
  \int_{t_0}^{t_1}
  \left(
    \|E_C(\varphi;t)\|\,\|O_L(t)\|
    +
    \|C_L(\varphi;t)\|\,\|E_O(t)\|
    +
    |E_K(\varphi;t)|
  \right)\dd t
  \label{eq:qcd-eec-endpoint-budget-bound}
\end{equation}
for any dual pair of finite vector norms.  In particular, exact operator
transport, exact coefficient transport, and the matching contact shift make
the retained endpoint observable independent of \(t\).
\end{proposition}

\begin{proof}
Differentiate \eqref{eq:qcd-eec-finite-endpoint-observable-chart} and insert
\eqref{eq:qcd-eec-endpoint-budget-operator}--\eqref{eq:qcd-eec-endpoint-budget-contact}:
\[
\begin{aligned}
  \partial_tV_L
  &=
  (C_L\gamma_{LL}+D_L+E_C)O_L
  +
  C_L(-\gamma_{LL}O_L+E_O)
  -
  D_LO_L
  +
  E_K .
\end{aligned}
\]
The anomalous-dimension terms cancel because the coefficient is a row and the
operator is a column.  The endpoint-extension row \(D_L\) cancels only because
the contact coordinate in
\eqref{eq:qcd-eec-endpoint-budget-contact} carries the compensating sign.
This gives \eqref{eq:qcd-eec-endpoint-budget-residual}.  Integrating in \(t\)
and applying the finite norm inequality for a row-column pairing gives
\eqref{eq:qcd-eec-endpoint-budget-bound}.
\end{proof}

The rows \(D_L\) are not new anomalous dimensions.  They record changes of
distributional representative at \(\rho=0\), such as the shift in
\eqref{eq:qcd-eec-plus-resolution-shift}.  The residuals \(E_O,E_C,E_K\)
are where an actual finite-basis endpoint calculation must place its omitted
operator channels, projected curvature, approximate loop-order truncation,
and endpoint matching defects.  A claimed all-order renormalized light-ray
OPE for the EEC must drive the right side of
\eqref{eq:qcd-eec-endpoint-budget-bound} to zero in the declared detector-test
topology, or must state the resulting approximation error.  Thus a mixing
matrix and a list of endpoint plus distributions become an observable theorem
only after this transport budget is supplied.

\paragraph{Minimal cusp-log flatness check.}
A useful first check is the part of a rapidity-factorized endpoint chart that
is forced by a cusp logarithm.  Work in a finite block in which the cusp
endomorphism \(A\) is fixed and commutes with itself at different scales.  Let
\(\Gamma_{\rm cusp}(t)\) denote the cusp coefficient in the same trace and
coupling convention as the endpoint calculation, and suppose the rapidity
dependence of the ultraviolet anomalous dimension contains
\begin{equation}
  \gamma_{\rm cusp}(t,r)
  =
  \Gamma_{\rm cusp}(t)\, r\, A .
  \label{eq:qcd-eec-cusp-log-gamma-chart}
\end{equation}
Then the rapidity anomalous dimension in the same chart must contain a term
\begin{equation}
  \eta_{\rm cusp}(t,r)
  =
  G_{\rm cusp}(t)\,A,
  \qquad
  \frac{\dd G_{\rm cusp}}{\dd t}=\Gamma_{\rm cusp}(t),
  \label{eq:qcd-eec-cusp-log-eta-chart}
\end{equation}
up to pieces whose \(t\)-derivative is zero in this block.  Indeed,
\[
  \partial_t\eta_{\rm cusp}
  =
  \Gamma_{\rm cusp}(t)A
  =
  \partial_r\gamma_{\rm cusp},
  \qquad
  [\gamma_{\rm cusp},\eta_{\rm cusp}]=0 .
\]
Thus the cusp-logarithmic part of
\eqref{eq:qcd-eec-light-ray-flatness} is satisfied.  With the opposite sign in
\eqref{eq:qcd-eec-cusp-log-eta-chart}, the curvature would be
\(-2\Gamma_{\rm cusp}(t)A\).  The sign statement belongs to the declared
operator convention \(\partial_t\mathbb O=-\gamma\mathbb O\),
\(\partial_r\mathbb O=-\eta\mathbb O\).  A full small-angle EEC calculation
must still compute the non-cusp matrices, endpoint contact counterterms, and
operator components beyond this cusp block in one declared rapidity scheme.

\begin{controlledapproximation}[Finite light-ray mixing truncation]
\label{ca:qcd-eec-finite-light-ray-mixing-truncation}
A finite small-angle EEC calculation based on a label set \(\mathcal I_L\)
is a controlled approximation only after it supplies:
\begin{enumerate}
  \item the regulated light-ray composites
  \(\mathbb O_I^{\rm reg}\) and the subtraction matrix \(Z\);
  \item the endpoint test-function space on which the coefficient
  distributions \(C_A(\rho)\), including plus distributions and contact
  coordinates, act;
  \item the matrices \(\gamma\) and, when present, \(\eta\), with the
  protected energy-moment row and its scheme convention;
  \item the transport matrices or path-ordered products solving
  \eqref{eq:qcd-eec-light-ray-transport-u}, and, in a two-scale scheme, the
  flatness condition \eqref{eq:qcd-eec-light-ray-flatness};
  \item a remainder functional \(R_L(\varphi,\Psi)\) for the omitted
  operators, stated in the topology of the matrix elements or cross-section
  distributions being computed.
\end{enumerate}
Without the last item, a finite light-ray basis is a useful coordinate
truncation but not an approximation theorem for the EEC distribution.
\end{controlledapproximation}

\begin{controlledapproximation}[Perturbative EEC endpoint calculations]
\label{ca:qcd-eec-endpoint-calculations}
A perturbative endpoint calculation of the EEC consists of a declared
endpoint variable, a factorization or light-ray OPE formula such as
\eqref{eq:qcd-eec-light-ray-ope-form} or
\eqref{eq:qcd-eec-back-to-back-factorization}, anomalous dimensions and
matching coefficients through specified orders, a prescription for matching
the endpoint expansion to the separated-angle fixed-order distribution, and a
power correction estimate in \(\theta\) or \(q_T/Q\).  The exact observable
remains the hadronic detector distribution
\eqref{eq:qcd-eec-definition}.
\end{controlledapproximation}

\paragraph{Endpoint matching as distributional gluing.}
The endpoint formulae above live in different distribution charts.  A
prediction for the full normalized EEC on \([-1,1]\) requires a gluing prescription,
because the contact coefficient at \(\zeta=1\), the perturbative
back-to-back delta coordinate at \(\zeta=-1\), and the separated-angle
density contribute to the same moment sum rules.

Let \(\mathcal R(\zeta)\) be the distribution on the open interval
\((-1,1)\) obtained from the declared bulk expression and the
overlap-subtracted small-angle and back-to-back endpoint formulae, all tested
against smooth functions whose support avoids the two endpoints.  Suppose its
first two moments exist after the declared endpoint subtractions:
\begin{equation}
  I_0[\mathcal R]=\langle \mathcal R,1\rangle,
  \qquad
  I_1[\mathcal R]=\langle \mathcal R,\zeta\rangle .
  \label{eq:qcd-eec-open-interval-moments}
\end{equation}
If the two endpoint atoms are retained as explicit scheme coordinates, write
\begin{equation}
  \mathcal E_{\rm match}(\zeta)
  =
  \mathcal R(\zeta)
  +K_+\,\delta(1-\zeta)
  +K_-\,\delta(1+\zeta).
  \label{eq:qcd-eec-matched-distribution-contact-terms}
\end{equation}
Here \(K_+\) is the coincident-detector contact coordinate and \(K_-\) is the
back-to-back endpoint coordinate generated by the perturbative source and
recoil expansion.  Only \(K_+\) has the direct eventwise interpretation
\eqref{eq:eventwise-eec-contact-weight}.  The two exact detector moment
identities \eqref{eq:eventwise-eec-first-sum-rules} impose
\begin{equation}
  I_0[\mathcal R]+K_+ +K_-=1,
  \qquad
  I_1[\mathcal R]+K_+ -K_-=0.
  \label{eq:qcd-eec-matching-moment-equations}
\end{equation}
When both endpoint atoms are left as matching coordinates, the solution is
\begin{equation}
  K_+
  =
  \frac{1-I_0[\mathcal R]-I_1[\mathcal R]}{2},
  \qquad
  K_-
  =
  \frac{1-I_0[\mathcal R]+I_1[\mathcal R]}{2}.
  \label{eq:qcd-eec-matching-delta-solution}
\end{equation}
If \(K_+\) is instead fixed independently by the detector contact definition,
then \eqref{eq:qcd-eec-matching-moment-equations} becomes a consistency test
on the bulk, small-angle, and back-to-back calculations in the chosen
subtraction scheme.  A mismatch is a statement about truncation, missing
power corrections, or incompatible endpoint conventions, rather than a new
observable.  This is the minimal matching structure required before a
perturbative endpoint formula can be compared with the nonperturbative
distribution \eqref{eq:qcd-eec-definition}.

\subsection{Multipoint Correlators and Energy-Flow Polynomials}

For a jet or event with a chosen positive energy scale \(E_{\mathrm{ref}}\),
define \(z_i=E_i/E_{\mathrm{ref}}\) and
\(\theta_{ij}\) to be the angular distance between directions \(i\) and \(j\).
For \(\beta>0\), the \(N\)-point energy correlation function used in jet
substructure has the form
\begin{equation}
  e_N^{(\beta)}
  =
  \sum_{i_1<\cdots<i_N}
  z_{i_1}\cdots z_{i_N}
  \prod_{a<b}
  \theta_{i_a i_b}^{\beta}.
  \label{eq:qcd-energy-correlation-function-jet}
\end{equation}
The strict positivity of \(\beta\) makes the observable collinear continuous:
when two labels become collinear, every factor distinguishing the split
vanishes or recombines into the total energy weight.  The soft behavior is
linear in the soft energy weight.  Thus \eqref{eq:qcd-energy-correlation-function-jet}
is infrared and collinear safe in the sense of the soft/collinear continuity
estimate for smeared energy flow above.

More generally, a finite graph \(G\) with vertices
\(\{1,\ldots,N\}\) defines an energy-flow polynomial
\begin{equation}
  \mathrm{EFP}_G^{(\beta)}
  =
  \sum_{i_1,\ldots,i_N}
  z_{i_1}\cdots z_{i_N}
  \prod_{(a,b)\in E(G)}
  \theta_{i_a i_b}^{\beta}.
  \label{eq:qcd-energy-flow-polynomial}
\end{equation}
This is again a polynomial functional of the calorimetric measure and angular
kernel.  It is a useful coordinate system on infrared-collinear safe event
information because it is built from the same positive detector measure
\(\mu_X\) rather than from colored partonic labels.  In an exact QCD
calculation the final states are hadrons; in a short-distance approximation,
quark and gluon variables enter through factorized coefficient functions,
evolution kernels, and power corrections for this color-singlet observable.

\subsection{Track Selectors and Charged-Energy Correlators}
\label{subsec:qcd-track-energy-correlators}

Collider detectors often measure a selected part of the calorimetric final
state, for example the energy carried by long-lived charged hadrons.  The
nonperturbative definition again begins with physical hadronic final states.
Let \(\mathfrak A_{\rm st}\) be the label set of stable hadron species in the
scattering representation, and fix a selector
\[
  \tau:\mathfrak A_{\rm st}\longrightarrow[0,1].
\]
For an ideal charged-track selector, \(\tau(a)=1\) for the species whose
charged tracks are included in the measurement and \(\tau(a)=0\) for the
species omitted by the tracking system.  A detector with efficiencies or
acceptance weights is represented by intermediate values of \(\tau\), once
those weights have been calibrated independently of the QCD short-distance
calculation.

For a hadronic final state
\[
  X=\{(p_r,a_r)\}_{r=1}^{N_X},
  \qquad
  \mathbf n_r=\mathbf p_r/|\mathbf p_r|,
\]
define the selected calorimetric measure
\begin{equation}
  \mu_X^\tau
  =
  \sum_{r\in X}\tau(a_r)p_r^0\,\delta_{\mathbf n_r}.
  \label{eq:qcd-track-calorimetric-measure}
\end{equation}
It satisfies \(0\le\mu_X^\tau\le\mu_X\) as positive measures.  The smeared
track detector on the outgoing multiplication representation is therefore
\[
  \widehat{\mathcal E}_X^\tau(f)
  =
  \int_{S^2}f\,\dd\mu_X^\tau
  =
  \sum_{r\in X}\tau(a_r)p_r^0 f(\mathbf n_r).
\]
Products of track detectors are products of this selected measure:
\begin{equation}
  \prod_{j=1}^{k}\widehat{\mathcal E}_X^\tau(f_j)
  =
  \int_{(S^2)^k}
  f_1(\mathbf n_1)\cdots f_k(\mathbf n_k)\,
  \dd\mu_X^\tau(\mathbf n_1)\cdots\dd\mu_X^\tau(\mathbf n_k).
  \label{eq:qcd-track-detector-product-measure}
\end{equation}
Thus track energy correlators are physical detector observables once the
selector \(\tau\) and the detector calibration have been specified.

The exact moment identities are selected-energy identities.  In a
center-of-mass event with \(z_r=p_r^0/Q\), set
\[
  e_\tau(X)=\sum_{r}\tau(a_r)z_r,\qquad
  \mathbf p_\tau(X)=\sum_{r}\tau(a_r)z_r\mathbf n_r .
\]
The eventwise two-point selected-energy distribution
\[
  E_{2,X}^{\tau}(\zeta)
  =
  \sum_{r,s\in X}
  \tau(a_r)\tau(a_s)z_rz_s\,
  \delta(\zeta-\mathbf n_r\cdot\mathbf n_s)
\]
obeys
\begin{equation}
  \int_{-1}^{1}E_{2,X}^{\tau}(\zeta)\,\dd\zeta
  =
  e_\tau(X)^2,
  \qquad
  \int_{-1}^{1}\zeta E_{2,X}^{\tau}(\zeta)\,\dd\zeta
  =
  |\mathbf p_\tau(X)|^2.
  \label{eq:qcd-track-eec-selected-moments}
\end{equation}
For \(\tau\equiv1\) these reduce to
\eqref{eq:eventwise-eec-first-sum-rules}.  For a proper track selector they
are additional nonperturbative event variables: the selected charged energy
and selected charged three-momentum fluctuate from event to event even when
the total energy and total three-momentum are fixed.

Perturbative calculations of selected-energy correlators require
nonperturbative track functions.  In a factorization scheme, a track function
for a parton label \(i\) is a positive distribution \(T_i(x,\mu)\) on
\([0,1]\), normalized by
\[
  \int_0^1 T_i(x,\mu)\,\dd x=1,
\]
whose moment
\[
  M_i^{(n)}(\mu)
  =
  \int_0^1 x^nT_i(x,\mu)\,\dd x
\]
is the \(n\)-th moment of the selected-energy fraction carried by hadrons in a
collinear jet initiated by \(i\), in the chosen scheme.  The functions
\(T_i\) are nonperturbative light-ray/fragmentation data; perturbation theory
controls their evolution and their convolution with hard and jet
coefficients.

The local collinear bookkeeping is fixed by energy partition.  If a parent
parton \(i\) splits into daughters \(j,k\) with energy fractions
\(z,1-z\), and the selected fractions in the daughter collinear sectors are
\(x_j,x_k\), then the selected fraction of the parent sector is
\[
  x=z x_j+(1-z)x_k .
\]
Therefore the real-splitting contribution to the \(n\)-th selected-fraction
moment has the binomial form
\begin{equation}
  M_{i\to jk}^{(n)}(z)
  =
  \sum_{a=0}^{n}
  \binom{n}{a}
  z^a(1-z)^{n-a}
  M_j^{(a)}M_k^{(n-a)} .
  \label{eq:qcd-track-function-binomial-moment}
\end{equation}
In particular,
\[
  M_{i\to jk}^{(1)}(z)=zM_j^{(1)}+(1-z)M_k^{(1)}
\]
and
\[
  M_{i\to jk}^{(2)}(z)
  =
  z^2M_j^{(2)}
  +(1-z)^2M_k^{(2)}
  +2z(1-z)M_j^{(1)}M_k^{(1)} .
\]
The last line separates the contact and resolved two-detector weights in a
track EEC calculation:
\begin{equation}
  w_{\tau,{\rm contact}}(z;j,k)
  =
  z^2M_j^{(2)}+(1-z)^2M_k^{(2)},\qquad
  w_{\tau,{\rm sep}}(z;j,k)
  =
  2z(1-z)M_j^{(1)}M_k^{(1)} .
  \label{eq:qcd-track-eec-contact-separated-weights}
\end{equation}
For the full calorimeter, \(M_i^{(n)}=1\) for every parton label and every
\(n\), and the separated weight reduces to the ordinary EEC weight
\(2z(1-z)\) in \eqref{eq:eec-ordered-collinear-weight}.

A controlled track-energy calculation must specify the selector \(\tau\), the
operator or fragmentation definition of the track functions \(T_i\), the
factorization scheme and scale \(\mu\), the evolution equation for \(T_i\) in
that scheme, the endpoint/contact convention for the selected detector
products, and the power remainder that compares the selected hadronic
observable \eqref{eq:qcd-track-detector-product-measure} with the finite
partonic factorization formula.  The selected-energy observable is defined at
the hadron level; the partonic track functions are the bridge used for its
short-distance expansion.

\subsection{Measured EEC Prediction Budgets}
\label{subsec:qcd-measured-eec-prediction-budgets}

The preceding endpoint and track formulae become physics only after they are
assembled into a measured final-state observable.  Let \(W_X(Q)\) denote the
normalized inclusive weight of the hadronic final state \(X\) produced by a
color-singlet source at center-of-mass energy \(Q\).  For a smooth angular
test \(\varphi\) on \([-1,1]\), the exact selected EEC observable is
\begin{equation}
  \mathcal O_\tau[\varphi;Q]
  =
  \sum_X W_X(Q)\,
  \int_{-1}^{1}
  \varphi(\zeta)\,E_{2,X}^{\tau}(\zeta)\,\dd\zeta .
  \label{eq:qcd-measured-selected-eec-observable}
\end{equation}
This is a color-singlet hadronic quantity.  The special case
\(\tau\equiv1\) is the full calorimetric EEC; a charged-track measurement is
obtained only after the selector and calibration entering
\(\mu_X^\tau\) have been fixed.

The exact selected moment constraints are
\begin{equation}
  \mathcal O_\tau[1;Q]
  =
  A_\tau(Q),
  \qquad
  \mathcal O_\tau[\zeta;Q]
  =
  B_\tau(Q),
  \label{eq:qcd-selected-eec-measured-moments}
\end{equation}
where
\[
  A_\tau(Q):=\sum_XW_X(Q)\,e_\tau(X)^2,
  \qquad
  B_\tau(Q):=\sum_XW_X(Q)\,|\mathbf p_\tau(X)|^2 .
\]
For the full calorimeter in a color-singlet center-of-mass event,
\(A_1(Q)=1\) and \(B_1(Q)=0\).  For a track selector these are
nonperturbative selected-energy moments.  They may be measured, modeled, or
factorized using track functions, but they are not fixed by total
four-momentum conservation alone.

Choose a smooth partition of the angular line into a bulk chart, a
small-angle chart near \(\zeta=1\), and a back-to-back chart near
\(\zeta=-1\).  In such a declared distributional chart, a leading-power
perturbative prediction has the form
\begin{equation}
\begin{aligned}
  \mathcal O_\tau[\varphi;Q]
  &=
  \mathcal P_{\tau}^{\rm bulk}[\varphi;Q,\mu]
  +
  \mathcal P_{\tau}^{\rm coll}[\varphi;Q,\mu,\nu]
  +
  \mathcal P_{\tau}^{\rm rec}[\varphi;Q,\mu,\nu]
  \\
  &\quad
  +
  K_{+}^{\tau}(Q)\,\varphi(1)
  +
  K_{-}^{\tau}(Q)\,\varphi(-1)
  +
  R_{\tau,Q}[\varphi].
\end{aligned}
\label{eq:qcd-measured-eec-factorization-budget}
\end{equation}
The three open-interval terms are understood to be overlap-subtracted in the
declared scheme.  The bulk term is the fixed-order separated-angle
coefficient.  The collinear term is the light-ray or jet endpoint expansion
near \(\zeta=1\), with the operator mixing, plus-distribution convention,
and contact coordinate discussed above.  The recoil term is the
back-to-back factorization chart, with its hard, jet, soft, rapidity, and
Sudakov data.  If \(\tau\ne1\), each partonic coefficient entering these
charts must be lifted by the track functions in the same factorization
scheme; diagonal detector slots use higher track moments as in
\eqref{eq:qcd-track-eec-contact-separated-weights}, not just first moments.

Let \(\mathcal P_{\tau}^{\rm open}\) be the sum of the three
overlap-subtracted open-chart terms in
\eqref{eq:qcd-measured-eec-factorization-budget}, and set
\[
  I_0^\tau=\mathcal P_{\tau}^{\rm open}[1],
  \qquad
  I_1^\tau=\mathcal P_{\tau}^{\rm open}[\zeta].
\]
If the endpoint atoms are retained as matching coordinates, the selected
moment constraints require
\begin{equation}
  I_0^\tau+K_+^\tau+K_-^\tau=A_\tau,
  \qquad
  I_1^\tau+K_+^\tau-K_-^\tau=B_\tau .
  \label{eq:qcd-selected-eec-endpoint-matching-equations}
\end{equation}
Thus
\begin{equation}
  K_+^\tau
  =
  \frac{A_\tau+B_\tau-I_0^\tau-I_1^\tau}{2},
  \qquad
  K_-^\tau
  =
  \frac{A_\tau-B_\tau-I_0^\tau+I_1^\tau}{2}.
  \label{eq:qcd-selected-eec-endpoint-atom-solution}
\end{equation}
For \(\tau\equiv1\) this reduces to
\eqref{eq:qcd-eec-matching-delta-solution}.  For a track measurement,
\(K_+^\tau\) contains the coincident selected-detector contact coordinate,
whereas \(K_-^\tau\) is a back-to-back endpoint coordinate of the matched
prediction.  Confusing these two endpoint atoms hides the physical difference
between selected-energy fragmentation data and recoil factorization.

\begin{controlledapproximation}[Measured EEC prediction]
\label{ca:qcd-measured-eec-prediction}
A controlled perturbative prediction for a calorimetric or selected-track EEC
must state:
\begin{enumerate}
  \item the hadronic detector observable
  \eqref{eq:qcd-measured-selected-eec-observable}, including the event
  normalization, selector \(\tau\), and detector calibration;
  \item the angular test space or finite bins, together with the endpoint
  partition and overlap-subtraction convention;
  \item the separated-angle fixed-order coefficient, the small-angle
  light-ray or jet endpoint chart, and the back-to-back recoil factorization
  chart, all in compatible subtraction schemes;
  \item the nonperturbative leading-power inputs, such as track functions or
  fragmentation data, and their evolution when the measurement is not fully
  calorimetric;
  \item the endpoint atoms \(K_\pm^\tau\), fixed or tested by
  \eqref{eq:qcd-selected-eec-endpoint-matching-equations};
  \item a residual functional \(R_{\tau,Q}\) separating perturbative
  truncation, omitted operator channels, endpoint matching, hadronization or
  shape-function effects, and ordinary power corrections.
\end{enumerate}
If the claimed errors in the three open charts are
\(\epsilon_{\rm bulk}(\varphi)\), \(\epsilon_{\rm coll}(\varphi)\), and
\(\epsilon_{\rm rec}(\varphi)\), if the endpoint atoms are known up to
\(\epsilon_+\) and \(\epsilon_-\), and if the remaining power or
hadronization error is \(\epsilon_{\rm pow}(\varphi)\), then the detector-test
pairing must obey a bound of the form
\begin{equation}
  \bigl|
    R_{\tau,Q}[\varphi]
  \bigr|
  \leq
  \epsilon_{\rm bulk}(\varphi)
  +
  \epsilon_{\rm coll}(\varphi)
  +
  \epsilon_{\rm rec}(\varphi)
  +
  |\varphi(1)|\epsilon_+
  +
  |\varphi(-1)|\epsilon_-
  +
  \epsilon_{\rm pow}(\varphi).
  \label{eq:qcd-measured-eec-residual-bound}
\end{equation}
Endpoint bins that probe \(Q\theta\) or \(q_T\) at hadronic scales are
outside a purely perturbative endpoint chart; the corresponding term in
\eqref{eq:qcd-measured-eec-residual-bound} must then be supplied by an
explicit nonperturbative shape, fragmentation, or track-function input.
\end{controlledapproximation}

\section{Current Correlators and the Hadronic Tensor}

Deep inelastic scattering probes a hadron by a local current.  The process is
inclusive:
\[
  e^-(k,\sigma)+N(P,s)\longrightarrow e^-(k',\sigma')+X,
  \qquad
  q:=k-k',
\]
where \(N\) is the target hadron and \(X\) runs over all hadronic final states
with total momentum \(P_X=P+q\).  The exchange momentum is spacelike,
\[
  Q^2:=-q^2>0,
  \qquad
  x_B:=\frac{Q^2}{2P\cdot q}.
\]
The variable \(x_B\) is the Bjorken scaling variable.

Figure~\ref{fig:qcd-dis-inclusive-kinematics} fixes the inclusive DIS
kinematics used below: the current transfers spacelike momentum \(q\), the
hadronic final state is summed over \(X\), and \(x_B=Q^2/(2P\cdot q)\).

\begin{figure}[H]
  \centering
  \resizebox{0.86\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize},
    photon/.style={decorate,decoration={snake,amplitude=1.35pt,segment length=6pt},
      line width=0.8pt,blue!70!black},
    fermion/.style={line width=0.85pt},
    hadron/.style={line width=0.9pt}]
    \coordinate (eIn) at (-4.3,1.45);
    \coordinate (eV) at (-0.8,1.05);
    \coordinate (eOut) at (3.0,1.72);
    \coordinate (hIn) at (-4.3,-1.2);
    \coordinate (hV) at (-0.8,-0.95);
    \coordinate (hOut) at (3.0,-0.55);

    \draw[fermion,-{Latex[length=2mm]}] (eIn)--(eV);
    \draw[fermion,-{Latex[length=2mm]}] (eV)--(eOut);
    \draw[photon,-{Latex[length=2mm]}] (eV)--(hV);
    \draw[hadron,-{Latex[length=2mm]}] (hIn)--(hV);
    \draw[hadron,-{Latex[length=2mm]}] (hV)--(hOut);
    \fill (eV) circle (2pt);
    \fill (hV) circle (2pt);

    \node[above] at (-3.1,1.45) {\(k,\sigma\)};
    \node[above] at (1.42,1.55) {\(k',\sigma'\)};
    \node[right,blue!70!black] at (-0.63,0.05) {\(q=k-k'\)};
    \node[below] at (-3.1,-1.22) {\(P,s\)};
    \node[below] at (1.40,-0.72) {\(X\)};
    \node[left] at (-4.35,1.45) {electron};
    \node[left] at (-4.35,-1.2) {hadron \(N\)};
    \node[right,align=left] at (3.08,-0.55) {summed hadronic\\final state};
  \end{tikzpicture}
  }
  \caption{Inclusive deep inelastic scattering.  The lepton line is treated
  perturbatively in the electromagnetic coupling, while the hadronic part is
  the current-current matrix element in QCD.}
  \label{fig:qcd-dis-inclusive-kinematics}
\end{figure}

Let the electromagnetic current be
\[
  J^\mu(x)=\sum_I e_I\,\bar q_I(x)\gamma^\mu q_I(x),
\]
where \(e_I\) is the electric charge of flavor \(I\).  At leading order in the
electromagnetic coupling, the amplitude for a fixed hadronic final state has
the form
\begin{equation}
  \mathcal A_{\sigma'\!,X;\sigma,s}
  =
  \bar u_{\sigma'}(k')\,e\gamma_\mu\,u_{\sigma}(k)\,
  \frac{-\ii}{q^2-\ii\epsilon}\,
  \bra{X}J^\mu(0)\ket{P,s}.
  \label{eq:qcd-dis-leading-amplitude}
\end{equation}
After summing over all hadronic final states \(X\) with the fixed transferred
momentum \(q\), the QCD part of the differential cross section is the
spin-averaged positive tensor
\begin{equation}
  W^{\mu\nu}(P,q)
  =
  \frac12\sum_s
  \sum_X (2\pi)^4\delta^{(4)}(P_X-P-q)\,
  \bra{P,s}J^\nu(0)\ket{X}
  \bra{X}J^\mu(0)\ket{P,s}.
  \label{eq:hadronic-tensor-definition}
\end{equation}
where \(\sum_X\) denotes the sum over species and the invariant phase-space
integrals for a complete set of hadronic states, using the same state
normalization as in the scattering chapters.  Completeness and translation
covariance give the equivalent form
\begin{equation}
  W^{\mu\nu}(P,q)
  =
  \frac12\sum_s
  \int\dd^4x\,\ee^{-\ii q\cdot x}
  \bra{P,s}J^\nu(x)J^\mu(0)\ket{P,s},
  \label{eq:hadronic-tensor-current-product}
\end{equation}
in those conventions.  Thus \(W^{\mu\nu}\) is a Wightman matrix element
between hadron states.  Its positivity follows directly from the sum over
\(X\) in
\eqref{eq:hadronic-tensor-definition}.

Current conservation gives
\[
  q_\mu W^{\mu\nu}=q_\nu W^{\mu\nu}=0 .
\]
For an unpolarized target, Lorentz covariance, current conservation, and
parity give two scalar structure functions.  Define
\[
  \Pi^\mu(P,q)
  :=
  P^\mu-\frac{P\cdot q}{q^2}q^\mu .
\]
Then
\begin{equation}
  W^{\mu\nu}(P,q)
  =
  \left(\eta^{\mu\nu}-\frac{q^\mu q^\nu}{q^2}\right)W_1(\nu,Q^2)
  +
  \Pi^\mu(P,q)\Pi^\nu(P,q)\,W_2(\nu,Q^2),
  \qquad
  \nu:=\frac{P\cdot q}{M},
  \label{eq:qcd-dis-tensor-decomposition}
\end{equation}
where \(M^2=P^2\).  Other common normalizations absorb powers of \(M\) and
\(P\cdot q\) into \(W_1,W_2\).

The support condition is kinematic.  If \(N\) is the lightest hadron with the
chosen conserved quantum numbers, then
\[
  (P+q)^2=P_X^2\ge M^2 .
\]
Since \(q^2=-Q^2\), this implies
\[
  2P\cdot q\ge Q^2,
  \qquad
  0<x_B\le1
\]
for the physical inclusive region.

The time-ordered forward Compton amplitude is
\begin{equation}
  T^{\mu\nu}(P,q)
  =
  \frac{\ii}{2}\sum_s
  \int\dd^4x\,\ee^{-\ii q\cdot x}
  \bra{P,s}\operatorname T\,J^\nu(x)J^\mu(0)\ket{P,s}.
  \label{eq:qcd-forward-compton-amplitude}
\end{equation}
Its discontinuity in the energy variable is determined by the same Wightman
matrix elements:
\begin{equation}
  \operatorname{Disc}\,T^{\mu\nu}(P,q)
  =
  2\pi
  \left[
    W^{\mu\nu}(P,q)
    +W^{\nu\mu}(P,-q)
  \right],
  \label{eq:qcd-dis-discontinuity}
\end{equation}
with the second term absent in the physical region \(P\cdot q>0\) for the
spacelike process.  Thus the inclusive cross section is controlled by the
absorptive part of a time-ordered current-current amplitude, and the
observable tensor is the Wightman object
\eqref{eq:hadronic-tensor-definition}.

The Bjorken limit is
\[
  Q^2\to\infty,
  \qquad
  x_B\ \text{fixed}.
\]
The short-distance region \(x\to0\) in
\eqref{eq:qcd-forward-compton-amplitude} is then controlled by the operator
product expansion of the two currents:
\begin{equation}
  \operatorname T\,J^\mu(x)J^\nu(0)
  \sim
  \sum_a C_a^{\mu\nu}(x,\mu)\,\mathcal O_a(0,\mu).
  \label{eq:dis-current-ope}
\end{equation}
The Wilson coefficients \(C_a^{\mu\nu}\) are determined by short-distance
renormalization and can be computed in perturbation theory at large \(Q\).  The
matrix elements
\(\bra{P}\mathcal O_a(0,\mu)\ket{P}\) are long-distance hadronic data.

\section{Twist, Parton Distributions, and Factorization}
\label{sec:qcd-twist-parton-distributions-factorization}

\subsection{Twist and Light-Ray Operator Definitions}
\label{subsec:qcd-twist-light-ray-operator-definitions}

For a local symmetric traceless operator of Lorentz spin \(s\) and scaling
dimension \(\Delta\), its twist is
\[
  \tau:=\Delta-s .
\]
In an interacting theory \(\Delta\) is a renormalized scaling exponent of the
chosen operator basis.  The word ``twist'' in this section is therefore a
statement about the renormalized leading-power expansion, not only about the
engineering dimension of a classical monomial.  In the Bjorken limit, the
leading terms in the OPE are the smallest-twist operators compatible with the
quantum numbers of the currents.  A convenient way to collect the tower of
leading-twist matrix elements is by light-ray operators.

Let \(n^\mu\) be a future-directed lightlike vector, \(n^2=0\), and choose the
normalization \(P\cdot n>0\) for the target momentum.  For a representation
\(R\) of the gauge group, define the straight Wilson transporter
\[
  W_R(\lambda n,0)
  =
  \mathcal P
  \exp\left\{\ii\int_0^\lambda
  n^\mu A_\mu^a(sn)t_R^a\,\dd s\right\}.
\]
Here \(A_\mu=A_\mu^at^a\) is the same connection coordinate used throughout the
chapter, with \(D_\mu=\partial_\mu-\ii A_\mu\) and
\(-\frac1{4g^2}\operatorname{tr}F_{\mu\nu}F^{\mu\nu}\).  The symbol
\(\mathcal P\) orders factors with larger parameter \(s\) to the left.

\paragraph{Gauge covariance of the open light-ray operator.}
Let \(U(x)\) be a smooth gauge transformation in the fundamental
representation, and let \(U_R(x)\) be its image in a representation \(R\).
With the transformation law
\[
  A_\mu\longmapsto U A_\mu U^{-1}-\ii(\partial_\mu U)U^{-1},
  \qquad
  q(x)\longmapsto U(x)q(x),
\]
the transporter satisfies
\begin{equation}
  W_R(\lambda n,0)
  \longmapsto
  U_R(\lambda n)W_R(\lambda n,0)U_R(0)^{-1}.
  \label{eq:qcd-wilson-transporter-gauge-transformation}
\end{equation}
Consequently
\[
  \bar q(\lambda n)W_{\square}(\lambda n,0)\gamma\cdot n\,q(0)
\]
is a color-singlet bilocal operator before renormalization, and its
renormalized descendants remain gauge invariant if the regularization and
subtraction scheme preserve background gauge covariance.

Let \(W_R(s,0)\) denote the partial transporter from \(0\) to \(sn\).  It is the
solution of
\[
  \frac{\dd}{\dd s}W_R(s,0)
  =
  \ii n^\mu A_\mu^a(sn)t_R^a W_R(s,0),
  \qquad
  W_R(0,0)=\mathbf 1 .
\]
After the gauge transformation, the product
\(U_R(sn)W_R(s,0)U_R(0)^{-1}\) obeys the same differential equation with the
transformed connection and has the same initial value at \(s=0\).  Uniqueness of
solutions to this finite-dimensional linear ordinary differential equation
gives the stated transformation law.  The endpoint factors cancel against the
transformations of \(\bar q(\lambda n)\) and \(q(0)\), because the gamma matrix
acts on spinor indices and is the identity in color space.

For an unpolarized target, write
\[
  \langle \mathcal X\rangle_{P,H}
  :=
  \frac12\sum_s\bra{P,s}\mathcal X\ket{P,s},
\]
with the one-particle normalization fixed in the scattering chapters.  The
renormalized quark light-ray operator and its parton distribution are
\begin{align}
  \mathbb O_q(\lambda n;\mu)
  &:=
  \left[
    \bar q(\lambda n)W_{\square}(\lambda n,0)\gamma\cdot n\,q(0)
  \right]_{\mu},
  \label{eq:qcd-quark-light-ray-operator}\\
  f_{q/H}(x,\mu)
  &:=
  \frac1{2P\cdot n}
  \int\frac{\dd\lambda}{2\pi}\,
  \ee^{-\ii x\lambda P\cdot n}
  \langle\mathbb O_q(\lambda n;\mu)\rangle_{P,H}.
  \label{eq:pdf-light-ray-definition}
\end{align}
The factor \(2P\cdot n\) fixes the convention so that a free quark target has
\(f_{q/q}(x)=\delta(1-x)\).  The antiquark distribution is defined by the
charge-conjugate bilocal with the fields interchanged and the transporter
running between the same endpoints.  The gluon distribution is defined from
the adjoint light-ray operator
\begin{align}
  \mathbb O_g(\lambda n;\mu)
  &:=
  \left[
    F^a_{n\rho}(\lambda n)
    W_{\rm ad}^{ab}(\lambda n,0)
    F_n^{b\rho}(0)
  \right]_{\mu},
  \qquad
  F_{n\rho}:=n^\mu F_{\mu\rho},
  \label{eq:qcd-gluon-light-ray-operator}\\
  f_{g/H}(x,\mu)
  &:=
  \frac1{2x(P\cdot n)^2}
  \int\frac{\dd\lambda}{2\pi}\,
  \ee^{-\ii x\lambda P\cdot n}
  \langle\mathbb O_g(\lambda n;\mu)\rangle_{P,H},
  \qquad 0<x<1 .
  \label{eq:qcd-gluon-pdf-definition}
\end{align}
The factor \(1/x\) is part of the conventional normalization of the gluon
light-ray distribution; it is what makes the \(N\)-th Mellin moment correspond
to the spin-\(N\) local gluon operator below.  Possible singular behavior at
\(x=0\) belongs to the small-\(x\) problem and is not hidden inside the
definition.

\begin{definition}[Renormalized integrated PDF coordinate]
\label{def:qcd-renormalized-integrated-pdf-datum}
For a hadron species \(H\), a renormalized integrated PDF coordinate consists of
the following objects.
\begin{enumerate}
  \item A nonperturbative QCD presentation, or a finite-regulator
  approximation to one, which supplies normalized color-singlet hadron states
  \(\ket{P,s}\) with \(P^2=M_H^2\) and \(P\cdot n>0\).
  \item A lightlike direction \(n^\mu\), an orientation of the straight
  segment from \(0\) to \(\lambda n\), and the representation labels of the
  Wilson transporters appearing in
  \eqref{eq:qcd-quark-light-ray-operator} and
  \eqref{eq:qcd-gluon-light-ray-operator}.
  \item A finite ultraviolet regulator \(\Lambda\) and regulated
  gauge-invariant bilocal composites
  \(\mathbb O_a^\Lambda(\lambda n)\), smeared in \(\lambda\) before any
  distributional limit is taken.
  \item A subtraction and factorization scheme at scale \(\mu\), encoded by
  a convolutional coordinate map \(Z_{ab}\) from
  \(\mathbb O_a^\Lambda\) to renormalized light-ray coordinates
  \(\mathbb O_b(\mu)\) in the sense of
  \eqref{eq:qcd-light-ray-renormalization-convolution}.
  \item A declared distribution space in the light-cone fraction variable
  \(x\), including the endpoint contact convention at \(x=0,1\), on which the
  Fourier transform in \(\lambda P\cdot n\) is paired with test functions.
\end{enumerate}
The integrated PDF \(f_{a/H}(x,\mu)\) is the resulting matrix-element
distribution.  In the conventions of
\eqref{eq:pdf-light-ray-definition} and
\eqref{eq:qcd-gluon-pdf-definition},
\begin{align}
  f_{q/H}(\,\cdot\,,\mu)
  &=
  \frac{1}{2P\cdot n}\,
  \mathcal F_{\lambda P\cdot n\to x}
  \bigl[\langle\mathbb O_q(\lambda n;\mu)\rangle_{P,H}\bigr],
  \label{eq:qcd-pdf-datum-quark-fourier-coordinate}\\
  f_{g/H}(\,\cdot\,,\mu)
  &=
  \frac{1}{2x(P\cdot n)^2}\,
  \mathcal F_{\lambda P\cdot n\to x}
  \bigl[\langle\mathbb O_g(\lambda n;\mu)\rangle_{P,H}\bigr],
  \label{eq:qcd-pdf-datum-gluon-fourier-coordinate}
\end{align}
where the second equality is an equality in the declared distribution space
on which multiplication by \(1/x\) has been specified.  The symbol
\(\mathcal F_{\lambda P\cdot n\to x}\) denotes the Fourier transform with
the phases displayed in the defining formulae above.  Equations
\eqref{eq:qcd-pdf-datum-quark-fourier-coordinate} and
\eqref{eq:qcd-pdf-datum-gluon-fourier-coordinate} define renormalized
operator coordinates.  Support in \(0\le x\le1\), positivity properties,
probability language, and the appearance of the same distribution in a
measured structure function are additional spectral, scheme, and
factorization statements; they are not part of the definition of the
renormalized light-ray matrix element.
\end{definition}

The preceding equations are nonperturbative operator definitions once the
underlying quantum field theory supplies the renormalized light-ray operators
\(\mathbb O_a(\lambda n;\mu)\) as operator-valued distributions, or at least as
matrix-element distributions between the hadron states under consideration.
They are not definitions in terms of on-shell quarks or gluons, and they do not
presuppose a colored particle interpretation.  In four-dimensional continuum
QCD this existence statement is tied to the still-open problem of constructing
the theory and to the additional ultraviolet problem of renormalizing lightlike
Wilson-line composites.  A gauge-invariant regulator can define finite
approximants, but the continuum lightcone limit, the operator-renormalization
limit, and the factorization theorem are separate assertions.  Equal-time
Euclidean bilocals used in lattice quasi-PDF or pseudo-PDF methods are therefore
matching observables for the light-ray definition, not the definition itself.

The subscript \(\mu\) in
\eqref{eq:qcd-quark-light-ray-operator}--\eqref{eq:qcd-gluon-pdf-definition}
means that the singular lightlike bilocal has been defined in a specified
regularization and factorization scheme.  Let
\(\mathbb O_a^\Lambda(x)\) denote the Fourier-coordinate operator obtained
from the same gauge-invariant light-ray composite in a finite ultraviolet
regulator \(\Lambda\), after the path geometry, endpoint prescription, and
finite regulator have been fixed.  This cutoff-regulated operator is not a
new physical observable; it is the finite-regulator coordinate from which the
renormalized light-ray coordinate is obtained.  For integrated collinear PDFs,
the coordinate change is a convolution in the light-cone momentum fraction:
\begin{equation}
  \mathbb O^{\Lambda}_{a}(x)
  =
  \sum_b
  \int_x^1\frac{\dd z}{z}\,
  Z_{ab}\!\left(\frac{x}{z},\mu,\Lambda\right)
  \mathbb O_b(z;\mu),
  \label{eq:qcd-light-ray-renormalization-convolution}
\end{equation}
where \(a,b\) range over quarks, antiquarks, and gluons, and the convolution
is understood after pairing with test functions in a distribution space whose
endpoint contact conventions have been declared.  This equation is a
definition of the renormalized light-ray coordinates.  It separates
regularization, the introduction of the finite object
\(\mathbb O_a^\Lambda\), from renormalization, the choice of the finite
coordinate map \(Z\) and of the operator coordinate \(\mathbb O_b(\mu)\).
For TMDs or other observables with unresolved transverse momentum, an
additional rapidity renormalization is required; it is not part of the
integrated PDF definition used in inclusive DIS.

The DGLAP kernel is the infinitesimal form of this operator-coordinate map,
not a transition probability for physical colored asymptotic states.  Let
\(Z^{-1}\) be the convolution inverse in the chosen finite channel basis.
Differentiating
\eqref{eq:qcd-light-ray-renormalization-convolution} at fixed
\(\mathbb O^\Lambda\) gives
\[
  0
  =
  \sum_b(\mu\partial_\mu Z_{ab})\otimes\mathbb O_b(\mu)
  +
  \sum_b Z_{ab}\otimes\mu\partial_\mu\mathbb O_b(\mu),
\]
where \(\otimes\) denotes the same light-cone fraction convolution.  Hence
\begin{equation}
  \mu\partial_\mu\mathbb O_c(x;\mu)
  =
  \sum_b
  \int_x^1\frac{\dd z}{z}\,
  P_{cb}\!\left(\frac{x}{z},g(\mu)\right)
  \mathbb O_b(z;\mu),
  \qquad
  P_{cb}:=
  -\sum_a Z^{-1}_{ca}\otimes\mu\partial_\mu Z_{ab}.
  \label{eq:qcd-light-ray-operator-dglap-rg}
\end{equation}
Taking hadron matrix elements of
\eqref{eq:qcd-light-ray-operator-dglap-rg} gives
\eqref{eq:dglap-evolution} below.  The finite perturbative calculation of
\(P\) is a calculation of the ultraviolet coordinate change of the light-ray
operator.  The subsequent use of \(P\) in a measured DIS observable still
requires the leading-power factorization statement.

\paragraph{Local moments of the light-ray definition.}
The light-ray definition contains the local twist-two tower as its Taylor
coefficients at zero separation.  This statement is often quoted as a formal
identity between PDFs and partons; here it is only a statement about the
renormalized Wilson-line operator in
\eqref{eq:qcd-quark-light-ray-operator}.  Define the left covariant
derivative on the transported antiquark endpoint by
\[
  \bar q\,\overleftarrow D_n
  :=
  n^\mu\partial_\mu\bar q+\ii\,\bar q\,n^\mu A_\mu,
  \qquad n^2=0 .
\]
The sign is fixed by the same convention
\(D_\mu q=(\partial_\mu-\ii A_\mu)q\) used in the Wilson transporter above.
Since
\[
  \frac{\dd}{\dd\lambda}
  \bigl(\bar q(\lambda n)W_\square(\lambda n,0)\bigr)
  =
  \bigl(\bar q\,\overleftarrow D_n\bigr)(\lambda n)
  W_\square(\lambda n,0),
\]
iteration gives the short-distance expansion
\begin{equation}
  \bar q(\lambda n)W_\square(\lambda n,0)\gamma\cdot n\,q(0)
  =
  \sum_{N\ge0}\frac{\lambda^N}{N!}\,
  \bar q(0)(\overleftarrow D_n)^N\gamma\cdot n\,q(0)
  \label{eq:qcd-quark-light-ray-taylor-left}
\end{equation}
as a formal operator identity before subtraction, and as a renormalized
operator-coordinate identity after the corresponding local and bilocal
subtractions have been matched.  It is useful to absorb the endpoint sign into
the moment operator
\begin{equation}
  \mathcal O_{q,N}^{\leftarrow}(n;\mu)
  :=
  \left[
    \bar q(0)(-\ii\overleftarrow D_n)^N
    \gamma\cdot n\,q(0)
  \right]_{\mu}.
  \label{eq:qcd-quark-left-moment-operator}
\end{equation}
The minus sign is not optional: a free one-quark matrix element of
\(\bar q(\lambda n)\gamma\cdot n q(0)\) is proportional to
\(\exp(\ii\lambda P\cdot n)\), so \(-\ii\overleftarrow D_n\) has eigenvalue
\(P\cdot n\) on the left endpoint.  With the Fourier convention in
\eqref{eq:pdf-light-ray-definition}, the inverse distributional relation is
\[
  \frac{1}{2P\cdot n}
  \langle\mathbb O_q(\lambda n;\mu)\rangle_{P,H}
  =
  \int_{\mathbb R}\dd x\,
  \ee^{\ii x\lambda P\cdot n}f_{q/H}(x,\mu),
\]
whenever the pairing with test functions justifies inversion.  Differentiating
at \(\lambda=0\) therefore gives, for \(N\ge0\),
\begin{equation}
  \int_0^1\dd x\,x^N f_{q/H}(x,\mu)
  =
  \frac{
    \langle\mathcal O_{q,N}^{\leftarrow}(n;\mu)\rangle_{P,H}
  }{
    2(P\cdot n)^{N+1}
  } .
  \label{eq:qcd-quark-pdf-moment-left-operator}
\end{equation}
The familiar symmetric traceless tensor
\(\bar q\gamma^{(\mu_0}\ii D^{\mu_1}\cdots\ii D^{\mu_N)}q\) is obtained by
polarizing in the null vector \(n\) and changing from the left-endpoint
representative to a symmetric representative.  In forward matrix elements the
difference is a sum of total-derivative representatives with fixed momentum
transfer zero; in off-forward matrix elements those total derivatives are
precisely the terms that produce GPD polynomiality in
Section~\ref{sec:qcd-gpd-light-ray-matrix-elements}.  Thus
\eqref{eq:qcd-quark-pdf-moment-left-operator} is the convention-clean moment
statement for the present forward PDF definition, while the symmetric local
tensor is a different coordinate on the same leading-twist operator tower.

For the gluon definition the extra factor \(1/x\) in
\eqref{eq:qcd-gluon-pdf-definition} shifts the moment index.  Let
\(\overleftarrow D_{n,{\rm ad}}\) be the left covariant derivative acting on
the adjoint field at \(\lambda n\), including the adjoint connection in the
same Wilson-transporter convention.  For \(N\ge1\), define
\begin{equation}
  \mathcal O_{g,N}^{\leftarrow}(n;\mu)
  :=
  \left[
    F^a_{n\rho}(0)
    (-\ii\overleftarrow D_{n,{\rm ad}})^{N-1}
    F_n^{a\rho}(0)
  \right]_{\mu},
  \label{eq:qcd-gluon-left-moment-operator}
\end{equation}
with the color contraction written after transporting the left endpoint to
zero.  The inverse relation following from
\eqref{eq:qcd-gluon-pdf-definition} is
\[
  \frac{1}{2(P\cdot n)^2}
  \langle\mathbb O_g(\lambda n;\mu)\rangle_{P,H}
  =
  \int_{\mathbb R}\dd x\,
  \ee^{\ii x\lambda P\cdot n}\,x f_{g/H}(x,\mu),
\]
and therefore
\begin{equation}
  \int_0^1\dd x\,x^N f_{g/H}(x,\mu)
  =
  \frac{
    \langle\mathcal O_{g,N}^{\leftarrow}(n;\mu)\rangle_{P,H}
  }{
    2(P\cdot n)^{N+1}
  },
  \qquad N\ge1 .
  \label{eq:qcd-gluon-pdf-moment-left-operator}
\end{equation}
Equations
\eqref{eq:qcd-quark-pdf-moment-left-operator} and
\eqref{eq:qcd-gluon-pdf-moment-left-operator} are moment identities inside a
specified renormalized operator scheme.  They do not assert positivity,
support, or factorization.  Positivity and support require the spectral
interpretation of the hadron matrix element in the relevant scattering
representation; factorization requires the separate leading-power hypothesis
stated below.

\subsection{Euclidean Spatial Bilocals and Large-Momentum Matching}
\label{subsec:qcd-euclidean-bilocals-quasi-pdfs}

Lattice access to the light-ray definitions
\eqref{eq:pdf-light-ray-definition}--\eqref{eq:qcd-gluon-pdf-definition}
uses different operator coordinates.  The basic object is a spacelike
equal-time bilocal, not a lightlike bilocal.  Let \(e_3\) be a fixed unit
spatial vector and let \(P^\mu=(E_P,0,0,P_3)\), \(P_3>0\), be a one-hadron
momentum in a finite-volume regulator with a transfer matrix and a
gauge-invariant Hilbert space.  For a Dirac matrix \(\Gamma\) and a real
spatial separation \(z\), define the regulated spatial Wilson-line operator
\begin{equation}
  \mathcal Q_\Gamma(z;\Lambda)
  :=
  \left[
    \bar q(z e_3)\,
    W_{\square}(z e_3,0)\,
    \Gamma\,q(0)
  \right]_{\Lambda}.
  \label{eq:qcd-spatial-bilocal-operator}
\end{equation}
Here \(\Lambda\) denotes the complete ultraviolet and finite-volume regulator
data, including the lattice spacing and box when the operator is realized on a
lattice.  The Wilson transporter is along the straight spatial segment.  Its
linear length divergence, endpoint divergences, and finite subtraction are
part of the operator-renormalization datum; they are not part of the
definition of the light-ray PDF.

At finite gauge-covariant regulator, the operator
\(\mathcal Q_\Gamma(z;\Lambda)\) is gauge invariant as a color-singlet
bilocal.  A background-gauge-covariant finite-part prescription therefore
produces a gauge-invariant renormalized spatial-bilocal matrix element.
This is the endpoint-cancellation argument of
the Wilson-transporter covariance calculation above, with the lightlike path
\(s n\) replaced by the spatial path \(s z e_3\), \(0\le s\le1\).  The
parallel transporter transforms as
\[
  W_{\square}(z e_3,0)
  \longmapsto
  U(z e_3)W_{\square}(z e_3,0)U(0)^{-1}.
\]
The factor \(U(0)^{-1}\) cancels the transformation of \(q(0)\), the factor
\(U(z e_3)\) cancels the transformation of \(\bar q(z e_3)\), and \(\Gamma\)
acts only on spinor indices.  The finite-part map changes regulator
coordinates and operator coordinates but carries no free color index when it
is background gauge covariant.

Let
\begin{equation}
  \mathcal M_{\Gamma/H}^{\mathcal R}(z,P_3)
  :=
  \frac{1}{2P_3}
  \bra{P,H}\mathcal Q_\Gamma^{\mathcal R}(z)\ket{P,H},
  \qquad
  \nu:=P_3z ,
  \label{eq:qcd-spatial-bilocal-matrix-element}
\end{equation}
where \(\mathcal R\) denotes a specified finite-part prescription.  A quasi-PDF
coordinate is the Fourier transform
\begin{equation}
  \widetilde f_{\Gamma/H}^{\mathcal R}(x,P_3)
  :=
  P_3\int_{\mathbb R}\frac{\dd z}{2\pi}\,
  \ee^{\ii xP_3z}\,
  \mathcal M_{\Gamma/H}^{\mathcal R}(z,P_3),
  \label{eq:qcd-quasi-pdf-coordinate}
\end{equation}
understood as a tempered distribution in \(x\) after smearing.  The prefactor
\(P_3\) fixes the inverse convention
\[
  \mathcal M_{\Gamma/H}^{\mathcal R}(z,P_3)
  =
  \int_{\mathbb R}\dd x\,
  \ee^{-\ii xP_3z}
  \widetilde f_{\Gamma/H}^{\mathcal R}(x,P_3)
\]
for distributions for which the Fourier inversion is valid.  A pseudo-PDF or
Ioffe-time coordinate instead regards the same matrix element as a function,
or distribution, of the Ioffe time \(\nu=P_3z\) and spacelike length
\(z^2>0\) in Euclidean notation:
\begin{equation}
  \mathfrak M_{\Gamma/H}^{\mathcal R}(\nu,z^2)
  :=
  \mathcal M_{\Gamma/H}^{\mathcal R}(z,P_3).
  \label{eq:qcd-pseudo-itd-coordinate}
\end{equation}
When \(\mathfrak M_{\Gamma/H}^{\mathcal R}(0,z^2)\ne0\), the reduced
Ioffe-time coordinate
\[
  \mathfrak M_{\rm red}(\nu,z^2)
  :=
  \frac{\mathfrak M_{\Gamma/H}^{\mathcal R}(\nu,z^2)}
       {\mathfrak M_{\Gamma/H}^{\mathcal R}(0,z^2)}
\]
cancels any multiplicative Wilson-line factor depending only on \(z^2\) in
that scheme.  It does not remove matching coefficients, target-mass effects,
or higher-twist matrix elements.

\begin{controlledapproximation}[Large-momentum matching statement]
\label{ca:qcd-large-momentum-quasi-pdf-matching}
Fix a renormalized spatial-bilocal scheme \(\mathcal R\), a light-ray
factorization scheme at scale \(\mu\), and a test function
\(h\in C_c^\infty(\mathbb R)\).  A large-momentum matching statement for
\eqref{eq:qcd-quasi-pdf-coordinate} consists of kernels
\(C_{\Gamma a}^{\mathcal R}\), with \(a\) running over the light-ray operator
basis, such that
\begin{align}
  \left\langle
  \widetilde f_{\Gamma/H}^{\mathcal R}(\,\cdot\,,P_3),h
  \right\rangle
  &=
  \sum_a\int_0^1\dd y\, f_{a/H}(y,\mu)
  \int_{\mathbb R}\dd x\,h(x)\,
  \frac1y
  C_{\Gamma a}^{\mathcal R}
  \left(\frac{x}{y},\frac{\mu}{P_3}\right)
  \notag\\
  &\hspace{1.0cm}
  +R_{\Gamma,H}^{\mathcal R}[h;P_3,\mu].
  \label{eq:qcd-quasi-pdf-matching-distributional}
\end{align}
The statement is controlled only after the topology of the remainder has been
specified.  A typical leading-power version requires, for fixed compact
support of \(h\),
\begin{equation}
  |R_{\Gamma,H}^{\mathcal R}[h;P_3,\mu]|
  \le
  C_h
  \left(
    \frac{M_H^2}{P_3^2}
    +\frac{\Lambda_{\rm had}^2}{P_3^2}
  \right)
  +
  R_{\rm reg}[h;a,L,P_3],
  \label{eq:qcd-quasi-pdf-power-remainder}
\end{equation}
where \(M_H\) is the hadron mass, \(\Lambda_{\rm had}\) is a hadronic scale
attached to higher-twist matrix elements, and \(R_{\rm reg}\) records finite
lattice-spacing, finite-volume, and finite-operator-length errors before the
continuum and infinite-volume limits are taken.  Endpoint regions, small
\(x\), and threshold \(x\to1\) require additional test-function restrictions
or separate factorization data.
\end{controlledapproximation}

The kernels in
Controlled Approximation~\ref{ca:qcd-large-momentum-quasi-pdf-matching} are
short-distance coordinates.  In perturbation theory they are determined by
matching matrix elements of the spatial bilocal and the light-ray operator in
the same infrared state, so the infrared dependence cancels in the difference.
That perturbative statement is not, by itself, a nonperturbative theorem about
the large-\(P_3\) limit of QCD.  A theorem-level version would need a
renormalized construction of the spatial Wilson-line operator, a spacelike
short-distance OPE with remainder estimates uniform after hadron matrix
elements, and a proof that the large-momentum limit converts the spacelike
operator coordinates to the light-ray coordinates in the stated distributional
topology.

\paragraph{Scheme covariance of finite matching kernels.}
Let \(f\) denote the column vector of light-ray PDFs in a convolution algebra,
and suppose a quasi-PDF coordinate satisfies
\(\widetilde f=C\otimes f\) at leading power.  If a finite light-ray scheme
change is \(f'=R\otimes f\), where \(R\) has a convolution inverse, then the
same quasi-PDF coordinate is written in the primed scheme as
\[
  \widetilde f=C'\otimes f',
  \qquad
  C'=C\otimes R^{-1}.
\]
If a linear charge functional \(N(f)=\sum_a n_a\int_0^1 f_a(y)\,\dd y\) is
preserved by \(R\) and the matching kernel satisfies
\[
  \sum_i m_i\int_{\mathbb R} C_{ia}(\xi)\,\dd\xi=n_a
\]
for the corresponding quasi-PDF charge weights \(m_i\), then the matched
quasi-PDF has the same charge \(N\) at leading power.

Associativity of convolution gives
\[
  (C\otimes R^{-1})\otimes(R\otimes f)
  =
  C\otimes(R^{-1}\otimes R)\otimes f
  =
  C\otimes f.
\]
For the charge statement, integrate the matching formula over \(x\) with
weights \(m_i\).  The change of variables \(x=y\xi\), \(y>0\), gives
\[
  \sum_i m_i\int_{\mathbb R}\dd x\,\frac1y C_{ia}(x/y)
  =
  \sum_i m_i\int_{\mathbb R}\dd\xi\,C_{ia}(\xi)
  =
  n_a.
\]
The remaining \(y\)-integral is exactly \(N(f)\).  The assumed preservation of
\(N\) by \(R\) says that the same value is obtained in the primed light-ray
coordinates.

\subsection{DIS Factorization and DGLAP Evolution}
\label{subsec:qcd-dis-factorization-dglap}

Figure~\ref{fig:qcd-dis-ope-light-ray-factorization} records the operator
factorization used for inclusive DIS: the short-distance current product is
matched to renormalized light-ray operators whose convolutional
renormalization gives DGLAP evolution.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small] at (0,2.35) {current product};
      \fill[blue!70!black] (-1.05,0.55) circle (2.4pt);
      \fill[blue!70!black] (1.05,0.55) circle (2.4pt);
      \draw[blue!70!black,thick] (-1.05,0.55)--(1.05,0.55);
      \node[above] at (-1.05,0.65) {\(J^\mu(x)\)};
      \node[above] at (1.05,0.65) {\(J^\nu(0)\)};
      \draw[<->,orange!85!black,thick] (-1.05,0.16)--(1.05,0.16);
      \node[orange!85!black,below] at (0,0.16) {\(|x|\sim Q^{-1}\)};
      \node[align=center] at (0,-0.8)
        {short-distance coefficients\\\(C_a^{\mu\nu}(x,\mu)\)};
    \end{scope}

    \draw[-{Latex[length=2mm]},thick] (2.25,0.45)--(3.75,0.45);
    \node[align=center] at (3.0,0.88) {OPE};

    \begin{scope}[shift={(5.35,0)}]
      \node[font=\small] at (0,2.35) {local tower};
      \draw[thick] (-1.4,0.85)--(1.4,0.85);
      \foreach \x in {-1.05,-0.55,-0.05,0.45,0.95} {
        \draw[blue!65!black,thick,-{Latex[length=1.6mm]}] (\x,0.15)--(\x,0.78);
      }
      \node[align=center] at (0,-0.68)
        {\(\bar q\,\gamma^{(\mu_1}\ii D^{\mu_2}\cdots \ii D^{\mu_N)}q\),\\
        \(F^{(\mu_1}{}_{\rho}\ii D^{\mu_2}\cdots F^{\mu_N)\rho}\)};
    \end{scope}

    \draw[-{Latex[length=2mm]},thick] (7.85,0.45)--(9.25,0.45);
    \node[align=center] at (8.55,0.88) {resum\\on light ray};

    \begin{scope}[shift={(11.0,0)}]
      \node[font=\small] at (0,2.35) {renormalized light-ray operator};
      \draw[purple!70!black,thick,decorate,
        decoration={snake,amplitude=0.85pt,segment length=7pt}]
        (-1.75,0.52)--(1.75,0.52);
      \fill[blue!70!black] (-1.75,0.52) circle (2.3pt);
      \fill[red!70!black] (1.75,0.52) circle (2.3pt);
      \node[below] at (-1.75,0.35) {\(\bar q(\lambda n)\)};
      \node[below] at (1.75,0.35) {\(q(0)\)};
      \node[purple!70!black,above] at (0,0.78) {\(W[\lambda n,0]\)};
      \node[align=center] at (0,-0.72)
        {matrix element in \(\ket P\)\\defines \(f_{a/H}(x,\mu)\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The short-distance current product is first expanded into local
  renormalized operators.  The leading-twist tower can then be packaged into
  gauge-invariant light-ray operators whose hadron matrix elements define
  parton distributions.  The Wilson segment \(W[\lambda n,0]\) is the open
  transporter connecting the colored fields in the bilocal operator.}
  \label{fig:qcd-dis-ope-light-ray-factorization}
\end{figure}

The factorization assertion has two logically separate parts.  The first is
the exact definition of the hadronic tensor as the current-current Wightman
matrix element \eqref{eq:hadronic-tensor-current-product}.  The second is the
leading-power short-distance statement that, after the current OPE and
light-ray renormalization have been performed, the same observable can be
coordinatized by universal hadronic light-ray matrix elements and perturbative
coefficient functions.

\paragraph{Common dependency ladder for QCD factorization.}
The local definitions above and the boundary sections below are used through a
single comparison architecture.  For a process family \(\mathfrak P\), a
theorem-level factorization statement has to provide:
\begin{enumerate}
  \item the exact color-singlet observable \(W_{\mathfrak P}\), usually a
  Wightman current tensor or a finite-regulator Euclidean matrix element;
  \item the gauge-invariant operator coordinates, including Wilson-line
  geometry and the ultraviolet, cusp, endpoint, or rapidity subtractions
  required by that geometry;
  \item the short-distance coefficient, impact-factor, or matching map in a
  stated distribution topology;
  \item the scale-transport equations and finite-scheme covariance making the
  auxiliary factorization coordinates cancel in the paired observable;
  \item the leading-region or boundary mechanism, including threshold,
  small-\(x\), soft, rapidity, \(Y\)-term, or Glauber items whenever the
  process probes them;
  \item a power-remainder estimate uniform on the declared test-function
  window;
  \item the continuum, infinite-volume, large-momentum, or rapidity-regulator
  limit when the preceding objects are first constructed at finite regulator.
\end{enumerate}
Equivalently, after pairing with a test function \(\varphi\) in the declared
topology, the comparison between the exact observable and the factorized
coordinate must be bounded by a sum of named errors,
\begin{equation}
\begin{aligned}
&
\left|
  \left\langle
    W_{\mathfrak P,Q}-\mathcal F_{\mathfrak P,Q},
    \varphi
  \right\rangle
\right|
\le
\mathcal E_{\mathfrak P,Q}^{\rm fact}(\varphi),
\\
&
\mathcal E_{\mathfrak P,Q}^{\rm fact}
:=
\epsilon_{\mathfrak P,Q}^{\rm obs}
+\epsilon_{\mathfrak P,Q}^{\rm op}
+\epsilon_{\mathfrak P,Q}^{\rm sub}
+\epsilon_{\mathfrak P,Q}^{\rm coeff}
+\epsilon_{\mathfrak P,Q}^{\rm evol}
+\epsilon_{\mathfrak P,Q}^{\rm bdry}
+\epsilon_{\mathfrak P,Q}^{\rm rem}
+\epsilon_{\mathfrak P,Q}^{\rm lim}.
\end{aligned}
\label{eq:qcd-common-factorization-budget}
\end{equation}
Here \(\mathcal F_{\mathfrak P,Q}\) denotes the claimed factorized coordinate.
For compact-\(x\) inclusive DIS, the boundary term records the exclusion of
threshold and small-\(x\) regions.  For Drell--Yan and other TMD observables it
contains the rapidity subtraction, the \(Y\)-term, and the Glauber status.  For
small-\(x\) observables it contains the Wilson-line rapidity state, the
projective generator limit, and the impact-factor map to a measured process.
The finite calculations retained in this chapter check individual entries in
this ladder; they are not substitutes for a proof that every required entry can
be made small on a common directed window.

\begin{hypothesis}[Leading-twist DIS factorization hypothesis]
\label{hyp:qcd-dis-leading-twist-factorization}
Fix a factorization scheme and a compact interval
\(K\Subset(0,1)\) of Bjorken \(x\).  A leading-twist factorization hypothesis for an
inclusive structure function \(F_i\) consists of coefficient distributions
\(C_{i,a}\), renormalized integrated PDFs \(f_{a/H}\), and constants
\(C_{K,\varphi}\), \(\Lambda_{\rm HT}\), and \(p>0\), such that for every test
function \(\varphi\in C_c^\infty(K)\) the difference between the exact
structure function and the factorized expression obeys
\[
  \left|
  \left\langle
  F_i-\sum_a C_{i,a}\otimes f_{a/H},\varphi
  \right\rangle
  \right|
  \le
  C_{K,\varphi}
  \left(\frac{\Lambda_{\rm HT}}{Q}\right)^p
\]
for all sufficiently large \(Q\).  The convolution is the one displayed in
\eqref{eq:dis-factorization}.  The compact-\(x\) restriction excludes the
separate threshold \(x\to1\) and small-\(x\) limits, whose large logarithms
require additional factorization data.
\end{hypothesis}

Under this hypothesis, factorization expresses a structure function \(F_i\) in
the form
\begin{equation}
  F_i(x_B,Q)
  =
  \sum_{a}
  \int_{x_B}^{1}\frac{\dd z}{z}\,
  C_{i,a}\!\left(\frac{x_B}{z},\frac{Q}{\mu},g(\mu)\right)
  f_{a/H}(z,\mu)
  +
  R_i(x_B,Q),
  \label{eq:dis-factorization}
\end{equation}
where \(a\) runs over quark, antiquark, and gluon parton species, and the
remainder \(R_i\) satisfies the distributional power bound in
Hypothesis~\ref{hyp:qcd-dis-leading-twist-factorization}.  The coefficients
\(C_{i,a}\) are short-distance kernels; the distributions \(f_{a/H}\) are
universal hadronic matrix elements for the chosen factorization scheme.  This
paragraph is not a statement about colored asymptotic particles.  The
``parton'' label \(a\) is a label in the light-ray operator basis and in the
coefficient algebra.

The role of the parton distributions in
\eqref{eq:dis-factorization} is also the point at which the nonabelian theory
departs from the abelian Bloch--Nordsieck construction of
Remark~\ref{rem:abelian-scope-bloch-nordsieck}.  A quark or gluon entering a
short-distance hard subprocess is a colored partonic variable, not a physical
asymptotic state.  Its collinear gluon cloud is tied to a color generator and
to the Wilson line in \eqref{eq:pdf-light-ray-definition}.  Summing over
unresolved final-state gluons gives the final-state part of an
infrared-collinear safe observable, but the singularities associated with a
massless incoming colored parton are absorbed into the renormalized light-ray
operator defining \(f_{a/H}\).  Equivalently, the collinear subtraction that
renormalizes the PDF is part of the definition of the factorized hadronic
observable.  The residual \(\mu\)-dependence of this subtraction is the
DGLAP evolution below.

The operator-coordinate RG equation
\eqref{eq:qcd-light-ray-operator-dglap-rg} gives the evolution equation
\begin{equation}
  \mu\frac{\dd}{\dd\mu} f_{a/H}(x,\mu)
  =
  \sum_b\int_x^1\frac{\dd z}{z}\,
  P_{ab}\!\left(\frac{x}{z},g(\mu)\right)f_{b/H}(z,\mu),
  \label{eq:dglap-evolution}
\end{equation}
where \(P_{ab}\) are splitting kernels, equivalently the anomalous-dimension
kernels of the renormalized light-ray operator coordinates in the sign
convention of
\eqref{eq:qcd-light-ray-operator-dglap-rg}.  This is the precise sense in
which short-distance quark and gluon variables control high-energy hadronic
observables while physical asymptotic states remain color singlets.  The
factorized observable is independent of the auxiliary scale only when the
coefficient functions obey the dual equation
\begin{equation}
  \mu\partial_\mu C_{i,b}
  =
  -\sum_a C_{i,a}\otimes P_{ab}.
  \label{eq:qcd-dis-coefficient-dual-rg}
\end{equation}
Then
\[
  \mu\partial_\mu
  \sum_a C_{i,a}\otimes f_{a/H}
  =
  \sum_b
  \left(
    \mu\partial_\mu C_{i,b}
    +
    \sum_a C_{i,a}\otimes P_{ab}
  \right)\otimes f_{b/H}
  =
  0
\]
inside the declared convolution algebra.  Any residual scale dependence in a
finite-order calculation is therefore a truncation or power-remainder effect,
not a change in the nonperturbative definition of the PDF.

\subsection{One-Loop Splitting Kernels and Sum-Rule Checks}
\label{subsec:qcd-one-loop-splitting-kernels-sum-rules}

At leading order, in the same convention as \eqref{eq:dglap-evolution},
\[
  P_{ab}(x,g)
  =
  \frac{g^2}{8\pi^2}P_{ab}^{(0)}(x)+O(g^4).
\]
For unpolarized distributions in the \(\overline{\mathrm{MS}}\) scheme,
\[
  C_F=\frac{N_c^2-1}{N_c},
  \qquad
  C_A=2N_c,
  \qquad
  T_R=1,
\]
in the trace normalization of \eqref{eq:qcd-casimir-definitions}.  These are
twice the standard half-trace numbers, while the overall perturbative
coefficient in \eqref{eq:dglap-evolution} is written in the corresponding
coupling coordinate.  Let
\begin{equation}
  D_0(x):=\left(\frac1{1-x}\right)_+
  \label{eq:qcd-d0-plus-distribution}
\end{equation}
be the plus distribution defined by
\[
  \int_0^1\dd x\,D_0(x)\,\varphi(x)
  =
  \int_0^1\dd x\,\frac{\varphi(x)-\varphi(1)}{1-x}
\]
for every smooth test function \(\varphi\).  Products such as \(xD_0(x)\) are
defined by multiplication of this distribution by the smooth function \(x\).
With this unambiguous convention, the nonzero leading kernels are
\begin{align}
  P_{q_iq_j}^{(0)}(x)
  &=
  \delta_{ij}\,
  C_F\left[
    2D_0(x)-1-x
    +\frac32\delta(1-x)
  \right],
  \notag\\
  P_{\bar q_i\bar q_j}^{(0)}(x)
  &=
  P_{q_iq_j}^{(0)}(x),
  \notag\\
  P_{q_i g}^{(0)}(x)
  &=
  P_{\bar q_i g}^{(0)}(x)
  =
  T_R\bigl[x^2+(1-x)^2\bigr],
  \notag\\
  P_{gq_i}^{(0)}(x)
  &=
  P_{g\bar q_i}^{(0)}(x)
  =
  C_F\,\frac{1+(1-x)^2}{x},
  \notag\\
  P_{gg}^{(0)}(x)
  &=
  2C_A\left[
    xD_0(x)
    +\frac{1-x}{x}
    +x(1-x)
  \right]
  +
  \frac{11C_A-4T_RN_f}{6}\delta(1-x).
  \label{eq:lo-dglap-splitting-kernels}
\end{align}
This notation avoids the common ambiguity between plus-prescribing the whole
rational function and plus-prescribing only the pole at \(x=1\).  For example,
\(2D_0-1-x\) is the distribution often abbreviated as
\((1+x^2)/(1-x)_+\) in particle-physics notation.  Equivalently, for the flavor
singlet quark distribution
\(\Sigma=\sum_{i=1}^{N_f}(q_i+\bar q_i)\),
\[
  \mu\frac{\dd}{\dd\mu}
  \begin{pmatrix}\Sigma\\ f_g\end{pmatrix}
  =
  \frac{g^2}{8\pi^2}
  \begin{pmatrix}
    P_{qq}^{(0)} & 2N_fP_{qg}^{(0)}\\
    P_{gq}^{(0)} & P_{gg}^{(0)}
  \end{pmatrix}
  \otimes
  \begin{pmatrix}\Sigma\\ f_g\end{pmatrix}
  +O(g^4),
\]
where \(\otimes\) denotes the convolution appearing in
\eqref{eq:dglap-evolution}.  Nonsinglet quark combinations evolve with the
same \(P_{qq}^{(0)}\) and do not mix with the gluon at this order.

The kernels in \eqref{eq:lo-dglap-splitting-kernels} obey the following
one-loop distributional sum-rule checks:
\[
  \int_0^1\dd x\,P_{qq}^{(0)}(x)=0,
\]
and the two momentum-column identities
\[
  \int_0^1\dd x\,x
  \bigl(P_{qq}^{(0)}(x)+P_{gq}^{(0)}(x)\bigr)=0,
  \qquad
  \int_0^1\dd x\,x
  \bigl(2N_fP_{qg}^{(0)}(x)+P_{gg}^{(0)}(x)\bigr)=0 .
\]
Thus nonsinglet quark number and the total light-cone momentum fraction are
preserved by the leading-order evolution.
For \(m\ge0\),
\[
  \int_0^1\dd x\,x^mD_0(x)
  =
  \int_0^1\dd x\,\frac{x^m-1}{1-x}
  =
  -H_m,
  \qquad
  H_m:=\sum_{r=1}^m\frac1r ,
\]
with \(H_0=0\).  The number sum rule is
\[
  C_F\left(2H_0-\int_0^1(1+x)\,\dd x+\frac32\right)=0.
\]
For a quark parent,
\[
  \int_0^1\dd x\,xP_{qq}^{(0)}(x)
  =
  C_F\left(-2-\frac12-\frac13+\frac32\right)
  =
  -\frac43C_F,
\]
while
\[
  \int_0^1\dd x\,xP_{gq}^{(0)}(x)
  =
  C_F\int_0^1\dd x\,\bigl[1+(1-x)^2\bigr]
  =
  \frac43C_F .
\]
For a gluon parent,
\[
  \int_0^1\dd x\,xP_{qg}^{(0)}(x)
  =
  T_R\int_0^1\dd x\,x\bigl[x^2+(1-x)^2\bigr]
  =
  \frac13T_R ,
\]
and
\[
  \int_0^1\dd x\,xP_{gg}^{(0)}(x)
  =
  2C_A\left(-H_2+\frac12+\frac1{12}\right)
  +\frac{11C_A-4T_RN_f}{6}
  =
  -\frac23T_RN_f .
\]
The two column sums therefore vanish.

\section{Transverse-Momentum Dependent Distributions and Rapidity Renormalization}
\label{sec:qcd-tmd-rapidity-renormalization}

Integrated PDFs are obtained after summing over all transverse momentum in the
collinear sector.  Observables differential in a transverse momentum, for
example color-singlet production with \(q_\perp^2\ll Q^2\), require a different
operator coordinate.  The relevant nonperturbative input is not the integrated
light-ray operator \eqref{eq:qcd-quark-light-ray-operator}; it is a
transverse-separated light-ray matrix element with Wilson lines whose geometry
creates rapidity divergences.  Rapidity subtraction is therefore part of the
definition, in the same sense that ultraviolet subtraction is part of the
definition of a renormalized local operator.

Let \(n,\bar n\) be future-directed lightlike vectors with
\(n\cdot\bar n=2\), and let \(b_\perp\) be transverse:
\[
  b_\perp\cdot n=b_\perp\cdot\bar n=0 .
\]
For a real parameter \(\lambda\), set
\[
  z^\mu=\lambda n^\mu+b_\perp^\mu .
\]
A finite rapidity regulator may be implemented by replacing lightlike Wilson
lines by Wilson lines along non-lightlike directions \(v,\bar v\), by an
analytic rapidity regulator, by a delta regulator, or by an equivalent
gauge-covariant subtraction.  The monograph denotes the regulator data
collectively by \(\rho\).  At finite \(\rho\) and finite ultraviolet cutoff
\(\Lambda\), let
\(\mathcal U^{[\eta]}_{v}(z,0;\rho,\Lambda)\) be a staple transporter from
\(0\) to \(z\), with longitudinal direction \(v\), transverse bridge
\(b_\perp\), and orientation \(\eta=+1\) for future-pointing and
\(\eta=-1\) for past-pointing staples.

\begin{definition}[TMD operator definition]
\label{def:qcd-tmd-operator-definition}
A quark TMD operator definition for a hadron \(H\) with momentum \(P\) consists of:
\begin{enumerate}
  \item the vectors \(n,\bar n\), a transverse separation \(b_\perp\), and an
  orientation \(\eta\) of the staple Wilson line;
  \item a finite rapidity regulator \(\rho\) and ultraviolet regulator
  \(\Lambda\), both chosen in a gauge-covariant way;
  \item the unsubtracted beam matrix element
  \[
    B^{\rm unsub}_{q/H}(x,b_\perp;\rho,\Lambda,\eta)
    =
    \frac1{2P\cdot n}
    \int\frac{\dd\lambda}{2\pi}\,
    \ee^{-\ii x\lambda P\cdot n}
    \Bigl\langle
    \bar q(z)\,
    \mathcal U^{[\eta]}_{v}(z,0;\rho,\Lambda)\,
    \gamma\cdot n\,q(0)
    \Bigr\rangle_{P,H};
  \]
  \item the vacuum soft factor
  \(S(b_\perp;\rho,\Lambda,\eta)\), defined by the trace of the closed Wilson
  geometry obtained by multiplying the corresponding \(v\)- and
  \(\bar v\)-staples in the vacuum;
  \item a finite-part prescription, denoted
  \(\operatorname{FP}_{\rho,\Lambda}\), which removes ultraviolet and rapidity
  regulator dependence and introduces the scales \(\mu\) and \(\zeta\).
\end{enumerate}
The renormalized TMD in this scheme is
\begin{equation}
  F_{q/H}^{[\eta]}(x,b_\perp;\mu,\zeta)
  =
  \operatorname{FP}_{\rho,\Lambda}
  \left[
    Z_q(\mu,\zeta;\rho,\Lambda)\,
    \frac{
      B^{\rm unsub}_{q/H}(x,b_\perp;\rho,\Lambda,\eta)
    }{
      S(b_\perp;\rho,\Lambda,\eta)^{1/2}
    }
  \right].
  \label{eq:qcd-renormalized-tmd-definition}
\end{equation}
The square root is a convention for assigning half of the soft subtraction to
each collinear sector.  A different finite assignment is a different TMD
scheme, related by a finite multiplicative transformation.  The variables
\(\mu\) and \(\zeta\) record, respectively, ultraviolet and rapidity
renormalization.  The orientation \(\eta\) is part of the operator definition;
time-reversal odd TMDs may change sign between future- and past-pointing
staples, as in semi-inclusive DIS versus Drell--Yan kinematics.
\end{definition}

At finite regulator, the numerator and soft factor in
\eqref{eq:qcd-renormalized-tmd-definition} are gauge invariant after taking
the color trace and endpoint fields into account.  Hence a background
gauge-covariant finite-part prescription produces a gauge-invariant
renormalized TMD matrix element.
Every Wilson segment transforms by multiplication with the gauge
transformation at its endpoints, as in
\eqref{eq:qcd-wilson-transporter-gauge-transformation}.  Along a staple, the
factors at internal junctions cancel between adjacent segments.  The remaining
endpoint factors at \(0\) and \(z\) cancel the gauge transformations of
\(q(0)\) and \(\bar q(z)\).  In the soft factor all Wilson lines are multiplied
inside a trace, so the residual base-point factor is removed by cyclicity of
the trace.  The finite-part map changes only regulator and scale coordinates;
if it is background gauge covariant, it cannot introduce an uncanceled color
index.

Rapidity evolution is the statement that changing the auxiliary rapidity
scale changes the coordinate description of the same physical transverse
momentum distribution.  It is most transparent in impact-parameter space,
where convolutions in transverse momentum become products.  Put
\[
  \ell_\mu:=\log\mu,
  \qquad
  \ell_\zeta:=\log\sqrt{\zeta}.
\]
For a fixed \(x\)-test function and fixed \(b_\perp\), suppose the paired TMD
matrix element is nonzero and obeys
\begin{align}
  \frac{\partial}{\partial\ell_\zeta}
  \log F_{q/H}^{[\eta]}(x,b_\perp;\mu,\zeta)
  &=
  \mathcal D_q(b_\perp;\mu),
  \label{eq:qcd-collins-soper-equation}\\
  \frac{\partial}{\partial\ell_\mu}
  \log F_{q/H}^{[\eta]}(x,b_\perp;\mu,\zeta)
  &=
  \gamma_F^q(g(\mu),\ell_\zeta-\ell_\mu).
  \label{eq:qcd-tmd-mu-evolution}
\end{align}
The first equation defines the Collins--Soper kernel
\(\mathcal D_q\) in the chosen convention.

\paragraph{Collins--Soper consistency condition.}
The two scale variables \(\ell_\mu\) and \(\ell_\zeta\) are coordinates on the
renormalized TMD scheme, so a differentiable TMD whose mixed scale derivatives
commute obeys the corresponding integrability condition.  Explicitly, if the
derivatives in
\eqref{eq:qcd-collins-soper-equation}--\eqref{eq:qcd-tmd-mu-evolution}
commute on the renormalized TMD distribution, then
\begin{equation}
  \frac{\partial}{\partial\ell_\mu}\mathcal D_q(b_\perp;\mu)
  =
  \frac{\partial}{\partial\ell_\zeta}
  \gamma_F^q(g(\mu),\ell_\zeta-\ell_\mu).
  \label{eq:qcd-collins-soper-integrability}
\end{equation}
In the convention
\[
  \gamma_F^q(g,\ell_\zeta-\ell_\mu)
  =
  \gamma_F^{q,0}(g)
  -2\Gamma_{\rm cusp}^q(g)(\ell_\zeta-\ell_\mu),
\]
this gives
\begin{equation}
  \frac{\partial}{\partial\log\mu}\mathcal D_q(b_\perp;\mu)
  =
  -2\Gamma_{\rm cusp}^q(g(\mu)).
  \label{eq:qcd-collins-soper-cusp}
\end{equation}
At small \(b_\perp\), the one-loop kernel compatible with the cusp
normalization of \eqref{eq:qcd-nonsinglet-large-n} is
\begin{equation}
  \mathcal D_q(b_\perp;\mu)
  =
  -\frac{g^2(\mu)C_F}{4\pi^2}
  \log\frac{\mu^2 b_\perp^2\ee^{2\gamma_{\rm E}}}{4}
  +O(g^4),
  \label{eq:qcd-collins-soper-kernel-one-loop}
\end{equation}
up to a finite \(b_\perp\)-dependent scheme term independent of \(\mu\).
The last display is a normalization check on the one-loop result:
\(\partial_{\log\mu}\log(\mu^2b_\perp^2\ee^{2\gamma_{\rm E}}/4)=2\), so
differentiating \eqref{eq:qcd-collins-soper-kernel-one-loop} gives
\(-g^2C_F/(2\pi^2)+O(g^4)\), which is
\(-2\Gamma_{\rm cusp}^q(g)+O(g^4)\) with
\(\Gamma_{\rm cusp}^q(g)=g^2C_F/(4\pi^2)+O(g^4)\) in the trace-delta
normalization.  The substantive QCD input is the regulated Wilson-line
operator definition, the existence and scheme dependence of rapidity
renormalization, and the perturbative or nonperturbative determination of
\(\Gamma_{\rm cusp}^q\) and \(\mathcal D_q\); the displayed consistency
relation only records the integrability of the chosen scale coordinates.

\begin{remark}[Finite TMD scheme changes]
\label{rem:qcd-tmd-finite-scheme-change}
Let \(R(b_\perp;\mu,\zeta)\) be a finite scalar function and set
\(F'=\ee^R F\).  Then
\[
  \mathcal D'=\mathcal D+\frac{\partial R}{\partial\ell_\zeta},
  \qquad
  \gamma_F'=\gamma_F+\frac{\partial R}{\partial\ell_\mu}.
\]
The consistency relation
\eqref{eq:qcd-collins-soper-integrability} is preserved.  Therefore the
separation of a rapidity-sensitive cross section into hard, TMD, and soft
coordinates is scheme dependent, whereas the final cross section is not.
\par
Take logarithms of \(F'=\ee^R F\) and differentiate with respect to
\(\ell_\zeta\) and \(\ell_\mu\).  The displayed transformations follow.  Taking
\(\partial_{\ell_\mu}\) of the first transformation and
\(\partial_{\ell_\zeta}\) of the second gives the same extra term
\(\partial_{\ell_\mu}\partial_{\ell_\zeta}R\).  Hence the difference between
the two sides of
\eqref{eq:qcd-collins-soper-integrability} is unchanged.
\end{remark}

\begin{hypothesis}[Small-\(q_\perp\) color-singlet TMD factorization hypothesis]
\label{hyp:qcd-tmd-factorization-hypothesis}
For a color-singlet final state of invariant mass \(Q\) produced in
hadron--hadron scattering, a leading-power small-\(q_\perp\) TMD
factorization hypothesis consists of TMD schemes for the two incoming hadrons, a
hard function \(H_{ij}(Q,\mu)\), rapidity scales satisfying
\(\zeta_A\zeta_B=Q^4\), a \(Y\)-term that matches the region
\(|q_\perp|\sim Q\), a statement about Glauber exchange, and a distributional
remainder estimate.  In the region
\(\Lambda_{\rm QCD}\lesssim |q_\perp|\ll Q\) the singular part has the
impact-parameter form
\begin{equation}
  \frac{\dd\sigma_{\rm sing}}{\dd Q^2\dd^2q_\perp}
  =
  \sum_{i,j}
  H_{ij}(Q,\mu)
  \int\frac{\dd^2 b_\perp}{(2\pi)^2}\,
  \ee^{\ii b_\perp\cdot q_\perp}
  F_{i/A}(x_A,b_\perp;\mu,\zeta_A)
  F_{j/B}(x_B,b_\perp;\mu,\zeta_B).
  \label{eq:qcd-tmd-factorization}
\end{equation}
The formula is a theorem only when the regulator construction, cancellation or
operator representation of Glauber exchange, and the topology of the
power-remainder estimate have all been supplied.
\end{hypothesis}

\begin{controlledapproximation}[Matched small-\(q_\perp\) TMD spectrum]
\label{ca:qcd-tmd-y-term-matching-window}
The TMD formula in
Hypothesis~\ref{hyp:qcd-tmd-factorization-hypothesis} controls only the
singular small-\(q_\perp\) coordinate.  A physical prediction on a tested
transverse-momentum spectrum also needs the fixed-order large-\(q_\perp\)
remainder.  Let \(\varphi(q_\perp)\) be a smooth compactly supported test
function in a color-singlet Drell--Yan or vector-boson window, and write
\[
  \mathfrak T_Q[\varphi]
  =
  \left\langle
    \frac{\dd\sigma}{\dd Q^2\dd^2q_\perp},
    \varphi
  \right\rangle
\]
for the exact tested spectrum in a finite regulator.  At fixed perturbative
order \(N\), let \(\sigma_{\rm FO}^{(N)}\) denote the ordinary fixed-order
coordinate and let
\((\sigma_{\rm sing}^{\rm TMD})_{[N]}\) be the expansion of the TMD singular
coordinate through the same order, in the same ultraviolet, rapidity, and
finite-subtraction scheme.  The \(Y\)-term is the tested distribution
\begin{equation}
  Y^{(N)}[\varphi]
  :=
  \left\langle
    \sigma_{\rm FO}^{(N)}
    -
    (\sigma_{\rm sing}^{\rm TMD})_{[N]},
    \varphi
  \right\rangle .
  \label{eq:qcd-tmd-y-term-tested-definition}
\end{equation}
The matched coordinate is therefore
\begin{equation}
  \mathcal M_{\rm TMD}^{(N)}[\varphi]
  :=
  \left\langle
    \sigma_{\rm sing}^{\rm TMD,resum},
    \varphi
  \right\rangle
  +
  Y^{(N)}[\varphi].
  \label{eq:qcd-tmd-matched-tested-spectrum}
\end{equation}
This subtraction is load-bearing: omitting
\((\sigma_{\rm sing}^{\rm TMD})_{[N]}\) counts the singular fixed-order terms
once in the resummed TMD coordinate and again inside \(\sigma_{\rm FO}^{(N)}\).

Rapidity scale cancellation is part of the same tested statement.  Parametrize
\(\zeta_A=Q^2\ee^{2\eta}\), \(\zeta_B=Q^2\ee^{-2\eta}\).  With the convention
\(\ell_\zeta=\log\sqrt{\zeta}\), this means
\[
  \ell_{\zeta_A}=\log Q+\eta,
  \qquad
  \ell_{\zeta_B}=\log Q-\eta .
\]
If the two incoming TMDs use the same Collins--Soper kernel in the
color-singlet channel and the hard factor has no residual
\(\eta\)-dependence, then
\begin{equation}
  \partial_\eta
  \log\!\left[
    H(Q,\mu)\,
    F_{i/A}(b_\perp;\mu,\zeta_A)\,
    F_{j/B}(b_\perp;\mu,\zeta_B)
  \right]
  =
  \mathcal D_i(b_\perp;\mu)-\mathcal D_j(b_\perp;\mu).
  \label{eq:qcd-tmd-rapidity-split-derivative}
\end{equation}
The factor is one, not two, because the Collins--Soper derivative above is with
respect to \(\log\sqrt{\zeta}\).  For the diagonal channel
\(\mathcal D_i=\mathcal D_j\) this vanishes.  In a mixing problem the
corresponding statement is a cancellation of rapidity anomalous-dimension
matrices in the paired hard/beam/soft coordinate.  If the kernels are
mismatched, if the product \(\zeta_A\zeta_B=Q^4\) is not held fixed, or if the
\(Y\)-term is evaluated in a different subtraction scheme, the residual
rapidity dependence is a matching error, not a new physical dependence of the
color-singlet spectrum.

A theorem-level matched TMD prediction is the assertion that
\begin{equation}
  \left|
  \mathfrak T_Q[\varphi]
  -
  \mathcal M_{\rm TMD}^{(N)}[\varphi]
  \right|
  \le
  \epsilon_{\rm TMD}[\varphi]
  +\epsilon_Y[\varphi]
  +\epsilon_{\rm G}[\varphi]
  +\epsilon_{\rm pwr}[\varphi]
  +\epsilon_{\rm reg}[\varphi],
  \label{eq:qcd-tmd-matched-spectrum-residual-bound}
\end{equation}
where the terms respectively record the singular-coordinate construction,
fixed-order \(Y\)-term truncation and matching, Glauber status, power
corrections such as \(q_\perp/Q\) and hadron masses, and regulator or
large-\(b_\perp\) residuals.  The finite Collins--Soper checks above verify only
the scale-coordinate algebra inside this statement; they do not prove the
existence of the exact QCD decomposition or the smallness of the residuals in
\eqref{eq:qcd-tmd-matched-spectrum-residual-bound}.
\end{controlledapproximation}

\section{Drell--Yan Factorization and Glauber Status}
\label{sec:qcd-drell-yan-glauber-status}

Drell--Yan production is the timelike current analogue of DIS:
\[
  h_A(P_A)+h_B(P_B)\longrightarrow \ell^+\ell^-+X,
  \qquad q=p_{\ell^+}+p_{\ell^-},
  \qquad q^2=Q^2>0 .
\]
The QCD object is again a Wightman current matrix element, now between a
two-hadron incoming state.  For unpolarized hadrons define
\begin{equation}
  W_{AB}^{\mu\nu}(P_A,P_B,q)
  =
  \frac1{N_{\rm spin}}
  \sum_{\sigma_A,\sigma_B}
  \sum_X
  (2\pi)^4\delta^{(4)}(P_A+P_B-q-P_X)
  \bra{P_A,P_B}J^\mu(0)\ket{X}
  \bra{X}J^\nu(0)\ket{P_A,P_B}.
  \label{eq:qcd-drell-yan-hadronic-tensor}
\end{equation}
Equivalently, by translation covariance and completeness,
\begin{equation}
  W_{AB}^{\mu\nu}(P_A,P_B,q)
  =
  \frac1{N_{\rm spin}}
  \sum_{\sigma_A,\sigma_B}
  \int\dd^4x\,\ee^{\ii q\cdot x}
  \bra{P_A,P_B}J^\mu(x)J^\nu(0)\ket{P_A,P_B},
  \label{eq:qcd-drell-yan-current-product}
\end{equation}
with the same state-normalization convention as in the scattering chapters.
The partonic formulae below are short-distance coordinates for this tensor;
they are not definitions of a cross section with colored asymptotic states.

\paragraph{Drell--Yan leading-power kinematics.}
Let
\[
  s:=(P_A+P_B)^2,
  \qquad
  \tau:=\frac{Q^2}{s},
\]
and choose the hadron center-of-mass frame with the \(A\)-hadron moving in the
positive longitudinal direction.  At leading power in hadron masses and
transverse momentum, the longitudinal momentum fractions entering the
short-distance hard collision are
\begin{equation}
  x_A=\sqrt{\tau}\,\ee^y,
  \qquad
  x_B=\sqrt{\tau}\,\ee^{-y},
  \label{eq:qcd-drell-yan-xa-xb}
\end{equation}
where
\[
  y:=\frac12\log\frac{q^0+q^3}{q^0-q^3}
\]
is the lepton-pair rapidity.  Hence
\[
  x_Ax_B=\tau,
  \qquad
  y=\frac12\log\frac{x_A}{x_B}.
\]

In light-cone coordinates adapted to the beam directions, write the leading
hadron momenta as
\[
  P_A^+=\sqrt{s},\quad P_A^-=0,
  \qquad
  P_B^-= \sqrt{s},\quad P_B^+=0,
\]
and neglect transverse components in the leading-power kinematic identity.
The active partons carry
\[
  q^+=x_AP_A^+,\qquad q^-=x_BP_B^-.
\]
Therefore \(Q^2=q^+q^-+O(q_\perp^2,M_{A,B}^2)=x_Ax_Bs+O(q_\perp^2,M_{A,B}^2)\),
which gives \(x_Ax_B=\tau\) at leading power.  The rapidity satisfies
\[
  y=\frac12\log\frac{q^+}{q^-}
  =
  \frac12\log\frac{x_A}{x_B}.
\]
Solving the two equations gives
\eqref{eq:qcd-drell-yan-xa-xb}.

\paragraph{Integrated collinear factorization coordinate.}
If the lepton-pair transverse momentum is not kept differential, the appropriate
leading-power coordinate is the integrated collinear factorization formula, not
the TMD formula below.  Fix a compact rapidity and invariant-mass window in
which \(x_A,x_B\in K\Subset(0,1)\) and the threshold and small-\(x\) boundaries
are excluded.  A leading-power integrated Drell--Yan factorization hypothesis
consists of renormalized integrated PDF coordinates
\(f_{i/A},f_{j/B}\), a short-distance coefficient distribution
\(C^{\rm DY}_{ij}\), and a distributional remainder bound such that
\begin{equation}
  \frac{\dd\sigma_{\rm DY}}{\dd Q^2\dd y}
  =
  \sum_{i,j}
  \int_{x_A}^{1}\frac{\dd \xi_A}{\xi_A}
  \int_{x_B}^{1}\frac{\dd \xi_B}{\xi_B}\,
  f_{i/A}(\xi_A,\mu)\,
  f_{j/B}(\xi_B,\mu)\,
  C^{\rm DY}_{ij}
  \!\left(\frac{x_A}{\xi_A},\frac{x_B}{\xi_B};
  Q,\mu\right)
  +
  R_{\rm DY}.
  \label{eq:qcd-drell-yan-integrated-factorization}
\end{equation}
The PDFs in \eqref{eq:qcd-drell-yan-integrated-factorization} are the
Wilson-line light-ray matrix elements of
Definition~\ref{def:qcd-renormalized-integrated-pdf-datum}.  The
coefficient \(C^{\rm DY}_{ij}\) is a hard matching distribution in the same
factorization scheme; it includes the chosen flux, electroweak-current, and
Jacobian normalization.  The remainder \(R_{\rm DY}\) contains the declared
power corrections, threshold and small-\(x\) exclusions, perturbative truncation
remainder, and the Glauber item discussed below.  Thus the formula is not an
identity between colored partonic cross sections.  It is a comparison between
the exact color-singlet Wightman tensor
\eqref{eq:qcd-drell-yan-current-product} and a factorized coordinate built from
two incoming light-ray operator matrix elements.

\begin{proposition}[Born Drell--Yan rapidity-bin coefficient]
\label{prop:qcd-drell-yan-born-rapidity-bin}
At leading power and Born order for a single virtual photon, the integrated
coefficient in
\eqref{eq:qcd-drell-yan-integrated-factorization} is supported at
\(z_A=z_B=1\).  With \(N_c\) colors and electromagnetic quark charge \(e_q\),
\begin{equation}
  C_{q\bar q}^{\rm DY,0}(z_A,z_B;Q)
  =
  C_{\bar q q}^{\rm DY,0}(z_A,z_B;Q)
  =
  \frac{4\pi\alpha_{\rm em}^2e_q^2}
       {3N_c\,Q^2s}\,
  \delta(1-z_A)\delta(1-z_B),
  \label{eq:qcd-drell-yan-born-coefficient}
\end{equation}
for the cross section differential in \(Q^2\) and \(y\), after integrating the
lepton-pair decay angles.  Consequently, for a compact test function
\(\varphi(Q^2,y)\) whose support keeps \(x_A,x_B\in(0,1)\),
\begin{equation}
\begin{aligned}
  \mathfrak T_{\rm DY}^{0}[\varphi]
  &=
  \sum_q
  \int\dd Q^2\,\dd y\,
  \varphi(Q^2,y)\,
  \frac{4\pi\alpha_{\rm em}^2e_q^2}
       {3N_c\,Q^2s}                                     \\
  &\qquad{}\times
  \Bigl[
    f_{q/A}(x_A,\mu)f_{\bar q/B}(x_B,\mu)
    +
    f_{\bar q/A}(x_A,\mu)f_{q/B}(x_B,\mu)
  \Bigr],
  \qquad
  x_{A,B}=\sqrt{\frac{Q^2}{s}}\,\ee^{\pm y}.
\end{aligned}
  \label{eq:qcd-drell-yan-born-tested-bin}
\end{equation}
\end{proposition}

\begin{proof}
The color-averaged partonic annihilation
\(q\bar q\to\gamma^\ast\to\ell^+\ell^-\) has the Born angular distribution
\[
  \frac{\dd\hat\sigma_{q\bar q}^{0}}{\dd\cos\theta}
  =
  \frac{\pi\alpha_{\rm em}^2e_q^2}{2N_cQ^2}
  (1+\cos^2\theta),
\]
where the factor \(1/N_c\) is the color average of the singlet electromagnetic
current.  Integrating \(\cos\theta\in[-1,1]\) gives
\[
  \hat\sigma_{q\bar q}^{0}(\hat s=Q^2)
  =
  \frac{4\pi\alpha_{\rm em}^2e_q^2}{3N_cQ^2}.
\]
At Born level the pair has zero transverse momentum, so
\[
  Q^2=s\xi_A\xi_B,
  \qquad
  y=\frac12\log\frac{\xi_A}{\xi_B}.
\]
The change of variables \((\xi_A,\xi_B)\mapsto(Q^2,y)\) has
\[
  \left|
  \det\frac{\partial(Q^2,y)}{\partial(\xi_A,\xi_B)}
  \right|=s,
\]
which supplies the factor \(1/s\) in
\eqref{eq:qcd-drell-yan-born-coefficient}.  The convolution measure in
\eqref{eq:qcd-drell-yan-integrated-factorization} is also part of the
normalization: for \(0<x<1\),
\[
  \int_x^1\frac{\dd\xi}{\xi}\,
  f(\xi)\,\delta\!\left(1-\frac{x}{\xi}\right)
  =
  f(x),
\]
because
\(\delta(1-x/\xi)=x\,\delta(\xi-x)\).  Inserting the two Born delta functions
therefore evaluates the two PDFs at \(x_A\) and \(x_B\), giving
\eqref{eq:qcd-drell-yan-born-tested-bin}.  This proposition instantiates the
factorized coordinate in one measured channel; the residuals of
Proposition~\ref{prop:qcd-drell-yan-tested-factorization-residual-budget}
remain the additional all-order QCD factorization data.
\end{proof}

The scale cancellation is the two-incoming-leg analogue of
\eqref{eq:qcd-dis-coefficient-dual-rg}.  In a Mellin or Laplace chart where the
two convolutions become products, let \(f_A(N_A,\mu)\) and \(f_B(N_B,\mu)\) be
column vectors in parton species, and write the factorized coordinate as
\begin{equation}
  \widetilde\Sigma_{\rm DY}(N_A,N_B;\mu)
  =
  f_A(N_A,\mu)^T
  \widetilde C_{\rm DY}(N_A,N_B;Q,\mu)
  f_B(N_B,\mu).
  \label{eq:qcd-drell-yan-transform-coordinate}
\end{equation}
If
\[
  \mu\frac{\dd f_A}{\dd\mu}=P_A f_A,
  \qquad
  \mu\frac{\dd f_B}{\dd\mu}=P_B f_B,
\]
then auxiliary-scale independence of
\eqref{eq:qcd-drell-yan-transform-coordinate} requires
\begin{equation}
  \mu\frac{\dd \widetilde C_{\rm DY}}{\dd\mu}
  =
  -P_A^T\widetilde C_{\rm DY}
  -\widetilde C_{\rm DY}P_B .
  \label{eq:qcd-drell-yan-coefficient-dual-rg}
\end{equation}
Indeed,
\[
  \mu\frac{\dd}{\dd\mu}
  \bigl(f_A^T\widetilde C_{\rm DY}f_B\bigr)
  =
  f_A^T
  \bigl(P_A^T\widetilde C_{\rm DY}
  +\mu\partial_\mu\widetilde C_{\rm DY}
  +\widetilde C_{\rm DY}P_B\bigr)
  f_B,
\]
which vanishes by
\eqref{eq:qcd-drell-yan-coefficient-dual-rg}.  A finite factorization-scheme
change \(f_A'=R_A f_A\), \(f_B'=R_B f_B\) preserves the same physical coordinate
only if
\[
  \widetilde C_{\rm DY}'=(R_A^{-1})^T\widetilde C_{\rm DY}R_B^{-1}.
\]
This matrix identity is the clean bookkeeping behind statements that the two
incoming collinear singularities have been absorbed into the PDFs.  A
factorization proof must still supply the leading-region analysis, soft
cancellation or soft factor, Glauber status, and the remainder topology; the
algebra above only fixes the paired RG structure once those objects exist.

\begin{controlledapproximation}[Integrated Drell--Yan theorem boundary]
\label{ca:qcd-drell-yan-integrated-theorem-boundary}
The exact object in integrated Drell--Yan is a tested distribution obtained
from the color-singlet Wightman tensor
\eqref{eq:qcd-drell-yan-current-product}, not the partonic coordinate
\eqref{eq:qcd-drell-yan-integrated-factorization}.  Fix a compact
\((Q^2,y)\)-window \(K\), away from threshold and small-\(x\) boundaries, and
let \(\mathcal D_K\) be the corresponding test-function space.  A
leading-power integrated factorization theorem for this window must supply:
\begin{enumerate}
  \item a finite-regulator definition of the exact tested functional
  \(\mathfrak T_{\rm DY}[\varphi]\), \(\varphi\in\mathcal D_K\), from
  \eqref{eq:qcd-drell-yan-current-product};
  \item the renormalized light-ray PDF coordinates and the hard coefficient in
  the same subtraction scheme;
  \item a leading-region map from the exact current product to the two
  incoming collinear sectors, with its error measured in
  \(\mathcal D_K^\ast\);
  \item a soft-cancellation or soft-factor construction in the same regulator;
  \item a Glauber statement strong enough for the chosen measurement topology,
  such as a commutator or inclusive-projection bound of the form described
  below;
  \item a power-remainder topology, including hadron-mass, threshold,
  small-\(x\), and perturbative-truncation qualifications;
  \item a regulator-limit statement identifying which residuals vanish and
  which remain as declared approximation errors.
\end{enumerate}
Thus the formula
\eqref{eq:qcd-drell-yan-integrated-factorization} becomes a theorem only after
the residuals in this list are bounded for the exact tested Wightman functional.
The paired RG equation \eqref{eq:qcd-drell-yan-coefficient-dual-rg} is one
coordinate-transport check inside this boundary; it is not the boundary itself.
\end{controlledapproximation}

\begin{proposition}[Tested Drell--Yan factorization residual budget]
\label{prop:qcd-drell-yan-tested-factorization-residual-budget}
In the setting of
Controlled Approximation~\ref{ca:qcd-drell-yan-integrated-theorem-boundary},
suppose the subtraction analysis represents the exact tested functional as
\begin{equation}
  \mathfrak T_{\rm DY}[\varphi]
  =
  f_A^T C_\varphi f_B
  +
  \Delta_{\rm lead}[\varphi]
  +
  \Delta_{\rm soft}[\varphi]
  +
  \Delta_{\rm G}[\varphi]
  +
  \Delta_{\rm pwr}[\varphi]
  +
  \Delta_{\rm reg}[\varphi],
  \label{eq:qcd-drell-yan-tested-residual-decomposition}
\end{equation}
where \(f_A^TC_\varphi f_B\) is the tested version of
\eqref{eq:qcd-drell-yan-integrated-factorization}.  Assume that, on a window
\(Q\ge Q_0\),
\begin{equation}
  |\Delta_a[\varphi]|
  \leq
  \epsilon_a(Q_0)\,\|\varphi\|_{\mathcal D_K},
  \qquad
  a\in\{\mathrm{lead},\mathrm{soft},\mathrm G,\mathrm{pwr},\mathrm{reg}\}.
  \label{eq:qcd-drell-yan-residual-component-bounds}
\end{equation}
Then
\begin{equation}
  \left|
  \mathfrak T_{\rm DY}[\varphi]-f_A^TC_\varphi f_B
  \right|
  \leq
  \left(
  \epsilon_{\rm lead}
  +\epsilon_{\rm soft}
  +\epsilon_{\rm G}
  +\epsilon_{\rm pwr}
  +\epsilon_{\rm reg}
  \right)
  \|\varphi\|_{\mathcal D_K}.
  \label{eq:qcd-drell-yan-tested-residual-budget-bound}
\end{equation}
If the paired PDF/coefficient transport
\eqref{eq:qcd-drell-yan-coefficient-dual-rg} is exact, then
\(f_A^TC_\varphi f_B\) is independent of the auxiliary factorization scale, but
the physical mismatch in
\eqref{eq:qcd-drell-yan-tested-residual-budget-bound} is controlled only by the
five residual norms above.  Scale covariance of the factorized coordinate does
not make any nonzero residual disappear.
\end{proposition}

\begin{proof}
Subtract the factorized coordinate from
\eqref{eq:qcd-drell-yan-tested-residual-decomposition} and apply the triangle
inequality together with
\eqref{eq:qcd-drell-yan-residual-component-bounds}.  The final statement is the
scale derivative calculation already displayed in
\eqref{eq:qcd-drell-yan-coefficient-dual-rg}: the derivative of
\(f_A^TC_\varphi f_B\) cancels between the two PDF legs and the dual hard
coefficient.  This cancellation concerns only the retained coordinate.  The
exact Wightman functional can agree with it in the declared topology only when
the listed residual functionals are separately bounded, or vanish in the
regulator and large-\(Q\) limits being claimed.
\end{proof}

\begin{hypothesis}[Drell--Yan small-\(q_\perp\) TMD factorization hypothesis]
\label{hyp:qcd-drell-yan-tmd-factorization-hypothesis}
A leading-power Drell--Yan TMD factorization hypothesis consists of:
\begin{enumerate}
  \item the renormalized current hard function \(H_{ij}(Q,\mu)\);
  \item past-pointing TMD operator data
  \(F_{i/A}^{[-]}(x_A,b_\perp;\mu,\zeta_A)\) and
  \(F_{j/B}^{[-]}(x_B,b_\perp;\mu,\zeta_B)\), in the notation of
  Definition~\ref{def:qcd-tmd-operator-definition};
  \item rapidity scales constrained by \(\zeta_A\zeta_B=Q^4\);
  \item a \(Y\)-term matching to the region \(|q_\perp|\sim Q\);
  \item a proof or controlled hypothesis for Glauber exchange;
  \item a distributional power-remainder estimate in \(q_\perp/Q\) and in the
  hadron-mass corrections.
\end{enumerate}
The singular small-\(q_\perp\) term is then represented as
\begin{equation}
  \frac{\dd\sigma_{\rm sing}}{\dd Q^2\dd y\,\dd^2q_\perp}
  =
  \sum_{i,j}
  H_{ij}(Q,\mu)
  \int\frac{\dd^2b_\perp}{(2\pi)^2}\,
  \ee^{\ii b_\perp\cdot q_\perp}
  F_{i/A}^{[-]}(x_A,b_\perp;\mu,\zeta_A)
  F_{j/B}^{[-]}(x_B,b_\perp;\mu,\zeta_B),
  \label{eq:qcd-drell-yan-tmd-factorization}
\end{equation}
plus the stated \(Y\)-term and power remainder.  The past-pointing staple
orientation is part of the operator definition.  It is the operator-level
source of the sign reversal of time-reversal-odd TMD coordinates between
semi-inclusive DIS and Drell--Yan, once the relevant antiunitary
transformation and spin conventions are specified.
\end{hypothesis}

\paragraph{Finite unitarity model for Glauber cancellation.}
Let \(\mathcal H_{\rm obs}\) be a finite-dimensional observed Hilbert space,
let \(\mathcal H_{\rm un}\) be an unobserved Hilbert space, let
\(\rho\) be a trace-class density matrix on
\(\mathcal H_{\rm obs}\otimes\mathcal H_{\rm un}\), and let \(M\) be an
observable acting only on \(\mathcal H_{\rm obs}\).  If the effect of a
Glauber exchange on the unobserved degrees of freedom is represented by a
unitary \(U_G\) acting only on \(\mathcal H_{\rm un}\), then
\begin{equation}
  \operatorname{Tr}_{\rm obs,un}
  \left[
    (M\otimes\mathbf1)
    (\mathbf1\otimes U_G)\rho(\mathbf1\otimes U_G^\dagger)
  \right]
  =
  \operatorname{Tr}_{\rm obs,un}
  \left[
    (M\otimes\mathbf1)\rho
  \right].
  \label{eq:qcd-finite-glauber-unitarity-identity}
\end{equation}

Use cyclicity of the full trace and
\((\mathbf1\otimes U_G^\dagger)(M\otimes\mathbf1)(\mathbf1\otimes U_G)
=M\otimes\mathbf1\).  The result is
\[
  \operatorname{Tr}
  \left[
    \rho
    (\mathbf1\otimes U_G^\dagger)(M\otimes\mathbf1)(\mathbf1\otimes U_G)
  \right]
  =
  \operatorname{Tr}\left[\rho(M\otimes\mathbf1)\right],
\]
which is
\eqref{eq:qcd-finite-glauber-unitarity-identity}.
Equivalently, the partial trace over the unobserved Hilbert space is invariant
under unitary conjugation on that same unobserved factor.  The finite model
therefore isolates the algebraic input used in inclusive Glauber-cancellation
arguments: the observed measurement must be insensitive to the unitary acting
on the summed-over sector.

The same finite model also gives the residual which must be bounded when the
measurement is not fully inclusive.  For a general tested measurement
\(M_\varphi\) on
\(\mathcal H_{\rm obs}\otimes\mathcal H_{\rm un}\), define
\begin{equation}
  \Delta_{\rm G}^{\rm fin}[\varphi]
  =
  \operatorname{Tr}
  \left[
    \rho
    \bigl(
      (\mathbf1\otimes U_G^\dagger)
      M_\varphi
      (\mathbf1\otimes U_G)
      -
      M_\varphi
    \bigr)
  \right].
  \label{eq:qcd-finite-glauber-resolved-measurement-residual}
\end{equation}
Then
\[
  \Delta_{\rm G}^{\rm fin}[\varphi]=0
  \quad\hbox{if}\quad
  [M_\varphi,\mathbf1\otimes U_G]=0,
\]
and in finite dimension
\begin{equation}
  |\Delta_{\rm G}^{\rm fin}[\varphi]|
  \leq
  \|\rho\|_1\,
  \left\|
      (\mathbf1\otimes U_G^\dagger)
      M_\varphi
      (\mathbf1\otimes U_G)
      -
      M_\varphi
  \right\|_{\rm op}.
  \label{eq:qcd-finite-glauber-resolved-measurement-bound}
\end{equation}
Thus the integrated Drell--Yan Glauber residual
\(\Delta_{\rm G}[\varphi]\) in
\eqref{eq:qcd-drell-yan-tested-residual-decomposition} is not a ceremonial
slot.  It asks whether, after the leading-region and contour-deformation
analysis has identified the regulated eikonal Glauber action, the tested
color-singlet observable is inclusive enough for the commutator to vanish, or
whether the remaining commutator is quantitatively small in the declared
measurement topology.

\begin{controlledapproximation}[What the finite Glauber calculation proves]
\label{ca:qcd-glauber-cancellation-status}
The finite unitarity calculation above is the algebraic identity that a
Drell--Yan factorization proof must realize after the leading regions, color
flow, rapidity regulator, and measurement topology have been fixed.  It does
not prove by itself that QCD Glauber exchanges act only on an unobserved
tensor factor, nor that the contour deformations needed to isolate the
leading regions are valid.  Those statements are part of the
Glauber item in Hypothesis~\ref{hyp:qcd-drell-yan-tmd-factorization-hypothesis}.  If the
observable measures hadronic radiation in a way that resolves the spectator
sector, or if a color exchange remains entangled with the measured hard
sector, the finite identity above no longer supplies the required
cancellation.
\end{controlledapproximation}

\section{Off-Forward Light-Ray Matrix Elements and GPDs}
\label{sec:qcd-gpd-light-ray-matrix-elements}

Generalized parton distributions are light-ray matrix elements between
different hadron momenta.  They do not require a rapidity subtraction in the
same way as TMDs, because the operator remains integrated over transverse
momentum.  Their new feature is off-forwardness.

Let \(p'\) and \(p\) be on-shell spin-\(\frac12\) hadron momenta, and define
\[
  P=\frac{p'+p}{2},
  \qquad
  \Delta=p'-p,
  \qquad
  t=\Delta^2,
  \qquad
  \xi=-\frac{\Delta\cdot n}{2P\cdot n}.
\]
For the symmetric light-ray separation
\[
  z^\mu=\lambda n^\mu ,
\]
define the quark vector GPDs \(H^q\) and \(E^q\) by
\begin{align}
  &\int\frac{\dd\lambda}{2\pi}\,
  \ee^{-\ii x\lambda P\cdot n}
  \bra{p',s'}
  \left[
  \bar q\!\left(\frac{\lambda n}{2}\right)
  W_{\square}\!\left(\frac{\lambda n}{2},-\frac{\lambda n}{2}\right)
  \gamma\cdot n\,
  q\!\left(-\frac{\lambda n}{2}\right)
  \right]_{\mu}
  \ket{p,s}
  \notag\\
  &\qquad =
  \bar u_{s'}(p')
  \left[
    \gamma\cdot n\,H^q(x,\xi,t;\mu)
    +\frac{\ii\sigma^{n\alpha}\Delta_\alpha}{2M}\,
    E^q(x,\xi,t;\mu)
  \right]
  u_s(p),
  \label{eq:qcd-gpd-definition}
\end{align}
where \(\sigma^{n\alpha}:=n_\mu\sigma^{\mu\alpha}\), and \(M\) is the hadron
mass.  The support and endpoint behavior of the distributions are properties
of the renormalized light-ray matrix elements and of the chosen hadronic
state; they are not imposed by a parton model.

\begin{lemma}[Polynomiality of GPD moments]
\label{lem:qcd-gpd-polynomiality}
Assume the renormalized twist-two local operators obtained from the light-ray
operator in \eqref{eq:qcd-gpd-definition} exist as local operator-valued
distributions between the hadron states under consideration.  Then the
\(N\)-th Mellin moments
\[
  \int_{-1}^{1}\dd x\,x^{N-1}H^q(x,\xi,t;\mu),
  \qquad
  \int_{-1}^{1}\dd x\,x^{N-1}E^q(x,\xi,t;\mu)
\]
are polynomials in \(\xi\) of degree at most \(N\), with coefficients that are
Lorentz-scalar form factors depending on \(t\) and \(\mu\).
\end{lemma}

\begin{proof}
Expanding the symmetric light-ray operator in \(\lambda\) gives the local
twist-two tower
\[
  \bar q\,\gamma^{(\mu_1}
  \ii D^{\mu_2}\cdots\ii D^{\mu_N)}q-\text{traces},
\]
contracted with \(n_{\mu_1}\cdots n_{\mu_N}\).  Since \(n^2=0\), all trace
terms vanish after contraction with these \(n\)'s.  Lorentz covariance of the
off-forward matrix element gives a finite decomposition into symmetric tensors
made from \(P\), \(\Delta\), the metric, and the spinor bilinears
\(\bar u(p')\gamma^\alpha u(p)\) and
\(\bar u(p')\ii\sigma^{\alpha\beta}\Delta_\beta u(p)\).  After contraction
with \(n_{\mu_1}\cdots n_{\mu_N}\), each term contains
\[
  (P\cdot n)^{N-k}(\Delta\cdot n)^k
  =
  (P\cdot n)^N(-2\xi)^k
\]
for some \(0\le k\le N\), multiplied by a scalar form factor depending only on
\(t\) and \(\mu\).  Dividing by the overall powers of \(P\cdot n\) in the GPD
decomposition therefore gives a polynomial in \(\xi\) of degree no larger than
\(N\).  This argument uses only locality, Lorentz covariance, and the
existence of the renormalized local twist-two tower; it does not use a
probability interpretation.
\end{proof}

Deeply virtual Compton scattering and related exclusive processes use these
GPDs in coefficient-function convolutions.  As in inclusive DIS and TMD
factorization, the GPD itself is the renormalized hadronic light-ray matrix
element.  The factorization theorem connecting it to a measured exclusive
amplitude is an additional leading-power assertion with its own endpoint,
Glauber, and power-remainder data.

\section{Asymptotic Freedom and Exclusive Pion Amplitudes}
\label{sec:qcd-exclusive-pion-asymptotic-freedom}

\subsection{Pion Light-Ray Data and Factorization Scheme}
\label{subsec:qcd-pion-light-ray-factorization-scheme}

Inclusive DIS uses forward light-ray matrix elements.  Exclusive pion
amplitudes use light-ray matrix elements between the vacuum and a one-pion
state.  This is the part of the pion \(S\)-matrix where asymptotic freedom
gives a direct large-momentum prediction: the short-distance subgraph is a
perturbatively renormalized quark-gluon amplitude, while the pion appears only
through a renormalized gauge-invariant light-ray matrix element.  The statement
is not that a pion is a weakly bound pair of on-shell quarks.  The pion input is
the same hadronic state defined by the nonperturbative QCD sector; only the
coefficient multiplying its light-ray matrix element is perturbative.

Let \(P^\mu\) be the momentum of a charged pion and let \(n^\mu\) be a
future-directed lightlike vector with \(P\cdot n>0\).  In the isospin-symmetric
two-flavor theory, define the charged-pion light-cone distribution amplitude
by the renormalized bilocal matrix element
\begin{equation}
  \bra{\vac}
  \left[
  \bar d(\lambda n)\,
  W_{\square}(\lambda n,0)\,
  \gamma\cdot n\,\gamma_5\,u(0)
  \right]_{\mu}
  \ket{\pi^+(P)}
  =
  \ii f_\pi\,P\cdot n
  \int_0^1\dd x\,
  \ee^{-\ii x\lambda P\cdot n}\,
  \varphi_\pi(x,\mu).
  \label{eq:qcd-pion-lcda-definition}
\end{equation}
The Wilson transporter is the same path-ordered fundamental transporter used
in \eqref{eq:qcd-quark-light-ray-operator}.  The normalization of \(f_\pi\) is
the axial-current normalization of Chapter~\ref{ch:volume-ii-21},
\[
  \bra{\vac}A_a^\mu(0)\ket{\pi^b(P)}
  =
  \ii f_\pi P^\mu\delta_a{}^b .
\]
Setting \(\lambda=0\) in
\eqref{eq:qcd-pion-lcda-definition} therefore fixes
\begin{equation}
  \int_0^1\dd x\,\varphi_\pi(x,\mu)=1 .
  \label{eq:qcd-pion-lcda-normalization}
\end{equation}
The support interval \([0,1]\) is a property of the leading-twist light-ray
operator between the vacuum and the one-pion state.  It is not a probability
postulate.  The function \(\varphi_\pi\) is a renormalized distribution in
\(x\), depending on the factorization scheme and scale.

\begin{definition}[Exclusive pion factorization scheme]
\label{def:qcd-exclusive-pion-factorization-scheme}
An exclusive large-\(Q\) pion factorization statement consists of:
\begin{enumerate}
  \item a renormalized QCD matrix element whose external hadrons and currents
  are gauge-invariant;
  \item a light-ray operator basis, here represented at leading twist by
  \(\varphi_\pi(x,\mu)\);
  \item a hard coefficient \(T_H\), computed with external collinear partons
  and subtracted in the same scheme as the light-ray operator;
  \item a convolution prescription in the momentum fractions;
  \item a topology for the remainder, for example
  \(O(\Lambda_{\rm QCD}^2/Q^2)\) after excluding endpoint regions by the
  stated hypotheses, together with the order in \(\alpha_s\) to which
  \(T_H\) is known.
\end{enumerate}
The scheme is part of the definition of the controlled approximation.  Without
it, a formula involving a pion distribution amplitude is only notation.
\end{definition}

\subsection{ERBL Evolution and Gegenbauer Coordinates}
\label{subsec:qcd-erbl-gegenbauer-coordinates}

The one-loop renormalization of the light-ray operator in
\eqref{eq:qcd-pion-lcda-definition} gives the Efremov--Radyushkin--Brodsky--
Lepage evolution equation
\begin{equation}
  \mu\frac{\dd}{\dd\mu}\varphi_\pi(x,\mu)
  =
  \frac{\alpha_s(\mu)}{2\pi}
  \int_0^1\dd y\,V(x,y)\varphi_\pi(y,\mu)
  +O(\alpha_s^2),
  \label{eq:qcd-erbl-evolution}
\end{equation}
where, in the trace-delta convention,
\begin{align}
  V(x,y)
  &=
  C_F
  \left[
  \theta(y-x)\frac{x}{y}
  \left(1+\frac1{y-x}\right)
  +
  \theta(x-y)\frac{1-x}{1-y}
  \left(1+\frac1{x-y}\right)
  \right]_+ .
  \label{eq:qcd-erbl-kernel}
\end{align}
The plus prescription acts on the \(x\)-variable:
\begin{equation}
  \int_0^1\dd x\,[K(x,y)]_+ f(x)
  :=
  \int_0^1\dd x\,K(x,y)\bigl(f(x)-f(y)\bigr),
  \label{eq:qcd-erbl-plus-prescription}
\end{equation}
for test functions smooth near \(x=y\).  Equation
\eqref{eq:qcd-erbl-plus-prescription} makes the conservation of the axial
current normalization manifest:
\[
  \int_0^1\dd x\,V(x,y)=0 .
\]

\begin{proposition}[Diagonalization of the leading ERBL kernel]
\label{prop:qcd-erbl-gegenbauer-diagonalization}
For \(n\ge0\), let \(C_n^{3/2}\) be the Gegenbauer polynomial orthogonal on
\([-1,1]\) with weight \((1-z^2)\).  The leading ERBL kernel preserves the
basis
\[
  \varphi_n(x):=6x(1-x)C_n^{3/2}(2x-1),
\]
and
\begin{equation}
  \int_0^1\dd y\,V(x,y)\varphi_n(y)
  =
  -\frac12\gamma_n^{(0)}\,\varphi_n(x),
  \label{eq:qcd-erbl-eigenvalue-equation}
\end{equation}
with
\begin{equation}
  \gamma_n^{(0)}
  =
  C_F\left[
  1-\frac{2}{(n+1)(n+2)}
  +4\sum_{k=2}^{n+1}\frac1k
  \right].
  \label{eq:qcd-erbl-anomalous-dimensions}
\end{equation}
In particular \(\gamma_0^{(0)}=0\) and
\(\gamma_n^{(0)}>0\) for every \(n\ge1\).
\end{proposition}

\begin{proof}
Write
\[
  P_n(x):=C_n^{3/2}(2x-1),\qquad
  \varphi_n(x)=6x(1-x)P_n(x).
\]
The plus prescription in \eqref{eq:qcd-erbl-plus-prescription} subtracts the
same raw kernel with the first argument integrated.  Therefore, for any
polynomial \(P\), the singular action of \eqref{eq:qcd-erbl-kernel} on
\(\varphi_P(y):=6y(1-y)P(y)\) is the following finite polynomial:
\begin{align}
  \frac1{C_F}\int_0^1\dd y\,V(x,y)\varphi_P(y)
  &=
  6(1-x)\int_0^x\dd y\,y\bigl(P(y)-P(x)\bigr)
     \left(1+\frac1{x-y}\right)
     \notag\\
  &\quad
  +6x\int_x^1\dd y\,(1-y)\bigl(P(y)-P(x)\bigr)
     \left(1+\frac1{y-x}\right).
  \label{eq:qcd-erbl-finite-polynomial-action}
\end{align}
This identity is only the definition of the plus distribution written at fixed
\(x\): the factors \(y(1-x)\) and \(x(1-y)\) come from
\((1-x)\varphi_P(y)/(1-y)\) and \(x\varphi_P(y)/y\), while the delta
subtraction supplies the terms with \(P(x)\).  The apparent poles are
removable because \(P(y)-P(x)\) is divisible by \(y-x\).

Let \(P(x)=\sum_{m=0}^N p_mx^m\), and put
\(H_m=\sum_{r=1}^m r^{-1}\), \(H_0=0\).  Evaluating the beta integrals in
\eqref{eq:qcd-erbl-finite-polynomial-action} gives the explicit coefficient
operator
\begin{align}
  \mathcal E[P](x)
  &:=
  \frac1{C_F}\int_0^1\dd y\,V(x,y)\varphi_P(y)
  \notag\\
  &=
  6(1-x)
  \left[
    \sum_{m=0}^N p_m\frac{x^{m+2}}{m+2}
    -\frac{x^2}{2}P(x)
    -\sum_{m=1}^N p_m x^{m+1}(H_{m+1}-1)
  \right]
  \notag\\
  &\quad
  +6x
  \left[
    \sum_{m=0}^N p_m
    \left(
      \frac1{(m+1)(m+2)}
      -\frac{x^{m+1}}{m+1}
      +\frac{x^{m+2}}{m+2}
    \right)
    -\frac{(1-x)^2}{2}P(x)
  \right.
  \notag\\
  &\hspace{7.2em}\left.
    +\sum_{m=1}^N p_m\sum_{j=0}^{m-1}x^j
    \left(
      \frac1{(m-j)(m-j+1)}
      -\frac{x^{m-j}}{m-j}
      +\frac{x^{m-j+1}}{m-j+1}
    \right)
  \right].
  \label{eq:qcd-erbl-coefficient-operator}
\end{align}
For \(P=P_n\), the coefficients are fixed by
\[
  P_0(x)=1,\qquad P_1(x)=3(2x-1),
\]
and by the Gegenbauer recurrence
\begin{equation}
  (n+1)P_{n+1}(x)
  =
  (2n+3)(2x-1)P_n(x)-(n+2)P_{n-1}(x).
  \label{eq:qcd-erbl-gegenbauer-recurrence}
\end{equation}
Substitution of this recurrence into the coefficient operator
\eqref{eq:qcd-erbl-coefficient-operator} gives, by induction on \(n\),
\begin{equation}
  \mathcal E[P_n](x)
  =
  -\frac12
  \left[
    1-\frac{2}{(n+1)(n+2)}
    +4\sum_{k=2}^{n+1}\frac1k
  \right]
  6x(1-x)P_n(x).
  \label{eq:qcd-erbl-polynomial-identity}
\end{equation}
The induction is finite algebra in the polynomial ring \(\mathbb Q[x]\):
\eqref{eq:qcd-erbl-coefficient-operator} gives the action on each monomial,
\eqref{eq:qcd-erbl-gegenbauer-recurrence} gives the next \(P_n\), and the
new harmonic term changes by \(H_{n+2}-H_{n+1}=1/(n+2)\).  The cases
\(n=0,1\) start the induction; \(n=0\) is the axial-current normalization
already expressed by \(\int_0^1\dd x\,V(x,y)=0\).  Restoring the overall
factor \(C_F\) proves \eqref{eq:qcd-erbl-eigenvalue-equation}.  The factor
\(1/2\) in \eqref{eq:qcd-erbl-eigenvalue-equation} is tied to the convention
that the evolution equation was written with \(\alpha_s/(2\pi)\) in
\eqref{eq:qcd-erbl-evolution}; in the common convention in which the moment
equation is written with \(\alpha_s/(4\pi)\), the same
\(\gamma_n^{(0)}\) appears without this extra displayed factor.
\end{proof}

The diagonal basis has a representation-theoretic origin: the collinear
conformal group acting on two light-ray quark fields has
\(SL(2,\mathbb R)\) quadratic Casimir diagonalized by
\(x(1-x)C_n^{3/2}(2x-1)\), and the one-loop kernel
\eqref{eq:qcd-erbl-kernel} is compatible with this action after the plus
subtraction.  The proof above keeps the normalization independent of that
interpretation by reducing the statement to the finite polynomial identity
\eqref{eq:qcd-erbl-polynomial-identity}.

Therefore a pion LCDA may be expanded as
\begin{equation}
  \varphi_\pi(x,\mu)
  =
  6x(1-x)
  \left[
  1+\sum_{n=1}^{\infty}a_n(\mu)C_n^{3/2}(2x-1)
  \right],
  \label{eq:qcd-pion-lcda-gegenbauer-expansion}
\end{equation}
with odd \(a_n\) vanishing in the isospin-symmetric charge-conjugation
channel.  At leading order
\begin{equation}
  \mu\frac{\dd}{\dd\mu}a_n(\mu)
  =
  -\frac{\alpha_s(\mu)}{4\pi}\gamma_n^{(0)}a_n(\mu)
  +O(\alpha_s^2),
  \label{eq:qcd-pion-gegenbauer-moment-rg}
\end{equation}
and asymptotic freedom drives all \(n\ge1\) moments to zero.  The ultraviolet
fixed shape is consequently
\begin{equation}
  \varphi_\pi^{\rm as}(x)=6x(1-x).
  \label{eq:qcd-pion-asymptotic-lcda}
\end{equation}
The approach to this shape is logarithmic and scheme dependent at finite
scale.  The statement \eqref{eq:qcd-pion-asymptotic-lcda} is therefore an
asymptotic consequence of the light-ray anomalous dimensions, not an empirical
model for the pion at moderate momentum transfer.

\subsection{Hard Exclusive Form Factors and Boundary Regimes}
\label{subsec:qcd-hard-exclusive-form-factors}

The spacelike charged-pion electromagnetic form factor is defined by
\begin{equation}
  \bra{\pi^+(P')}J^\mu_{\rm em}(0)\ket{\pi^+(P)}
  =
  (P+P')^\mu F_\pi(Q^2),
  \qquad
  Q^2:=-(P'-P)^2>0 .
  \label{eq:qcd-pion-em-form-factor-definition}
\end{equation}
At leading twist and leading order in the hard coefficient,
\begin{equation}
  F_\pi(Q^2)
  =
  \frac{4\pi\alpha_s(\mu)C_F}{N_c}\,
  \frac{f_\pi^2}{Q^2}
  \left[
  \int_0^1\dd x\,\frac{\varphi_\pi(x,\mu)}{x}
  \right]
  \left[
  \int_0^1\dd y\,\frac{\varphi_\pi(y,\mu)}{y}
  \right]
  +O(\alpha_s^2,Q^{-4})_{\rm fac}.
  \label{eq:qcd-pion-form-factor-factorization}
\end{equation}
The subscript ``fac'' means that the power remainder belongs to the
factorization scheme of Definition~\ref{def:qcd-exclusive-pion-factorization-scheme};
endpoint sensitivity can modify the practical convergence at accessible
\(Q^2\) unless controlled by an additional Sudakov or nonperturbative input.

\paragraph{Derivation of the leading hard coefficient.}
Project each pion onto the color-singlet leading-twist collinear pair defined
by \eqref{eq:qcd-pion-lcda-definition}.  A single hard gluon must transfer
large transverse momentum between the two collinear quark lines; without this
exchange the electromagnetic current cannot convert the initial collinear
color-singlet into the final one at fixed large \(Q^2\).  The tree hard
subgraph contains one gluon propagator of virtuality \(O(Q^2)\) and one
internal quark propagator of virtuality \(O(Q^2)\).  Contracting the spin
projectors fixed by the axial matrix element gives the numerator
proportional to \(P\cdot P'\), and the color-singlet projection gives
\[
  \frac1{N_c}\,t^a_{ij}t^a_{ji}
  =
  \frac{C_F}{N_c}.
\]
The two endpoint denominators are the light-cone momentum fractions carried by
the active quarks.  With the LCDA normalization
\eqref{eq:qcd-pion-lcda-normalization}, the sum of the two photon attachments
to the charged pion gives
\eqref{eq:qcd-pion-form-factor-factorization}.  No on-shell colored external
state appears in the final formula; colored partons were only projectors used
to compute the short-distance coefficient in the chosen factorization scheme.

Substituting \(\varphi_\pi^{\rm as}\) gives
\begin{equation}
  Q^2F_\pi(Q^2)
  \sim
  \frac{36\pi\alpha_s(Q)C_F}{N_c}\,f_\pi^2 .
  \label{eq:qcd-pion-form-factor-asymptotic-general}
\end{equation}
For \(N_c=3\), this is
\[
  Q^2F_\pi(Q^2)\sim 16\pi\alpha_{\rm ht}(Q) f_\pi^2
\]
in the common half-trace convention.  Equivalently, if one uses the
phenomenological convention \(f_\pi^{(130)}=\sqrt2\,f_\pi\), the same statement
is often written as \(Q^2F_\pi\sim
8\pi\alpha_{\rm ht}(f_\pi^{(130)})^2\).  The monograph keeps the
Chapter~\ref{ch:volume-ii-21} convention \(f_\pi\) fixed, so
\eqref{eq:qcd-pion-form-factor-asymptotic-general} is the convention-safe
formula.

The neutral-pion transition form factor is defined by
\begin{equation}
  \ii\int\dd^4x\,\ee^{\ii q\cdot x}
  \bra{\pi^0(P)}
  \operatorname T\,J^\mu_{\rm em}(x)J^\nu_{\rm em}(0)
  \ket{\vac}
  =
  \epsilon^{\mu\nu\rho\sigma}q_\rho k_\sigma\,
  F_{\pi^0\gamma\gamma^*}(Q^2),
  \label{eq:qcd-pion-transition-form-factor-definition}
\end{equation}
with \(k^2=0\), \(q^2=-Q^2\), and \(P=q+k\).  The low-energy endpoint is fixed
by the anomaly calculation of Chapters~\ref{ch:volume-ii-20} and
\ref{ch:volume-ii-21}:
\begin{equation}
  F_{\pi^0\gamma\gamma^*}(0)=\frac1{4\pi^2f_\pi}
  \label{eq:qcd-pion-transition-anomaly-normalization}
\end{equation}
in the present \(f_\pi\) convention.  At large spacelike \(Q^2\), leading
twist factorization gives
\begin{equation}
  F_{\pi^0\gamma\gamma^*}(Q^2)
  =
  \frac{f_\pi}{3Q^2}
  \int_0^1\dd x\,\varphi_\pi(x,\mu)
  \left(\frac1x+\frac1{1-x}\right)
  +O(\alpha_s,Q^{-4})_{\rm fac}.
  \label{eq:qcd-pion-transition-factorization}
\end{equation}
For the asymptotic LCDA,
\begin{equation}
  Q^2F_{\pi^0\gamma\gamma^*}(Q^2)\longrightarrow 2f_\pi .
  \label{eq:qcd-pion-transition-asymptotic}
\end{equation}
In the \(f_\pi^{(130)}=\sqrt2 f_\pi\) convention this is
\(\sqrt2 f_\pi^{(130)}\).  The anomaly value
\eqref{eq:qcd-pion-transition-anomaly-normalization} and the asymptotic value
\eqref{eq:qcd-pion-transition-asymptotic} constrain two different boundary
regimes of the same form factor.  They do not imply a unique interpolation.
Published high-\(Q^2\) measurements are therefore tests of the approach to the
factorized regime and of LCDA shape assumptions, not independent definitions
of the anomaly or of asymptotic freedom.

The BaBar/Belle tension in the neutral-pion transition form factor is best
formulated in exactly these terms.  A reported finite-\(Q^2\) trend can
challenge a chosen LCDA model, a treatment of endpoint regions, or an estimate
of power corrections.  It cannot by itself refute
\eqref{eq:qcd-pion-transition-anomaly-normalization} or
\eqref{eq:qcd-pion-transition-asymptotic}, because those two equations live at
opposite ends of the form factor and rest on different controlled arguments:
the first on the axial anomaly and the second on leading-twist factorization
plus ERBL evolution.  Any quantitative comparison to data must therefore state
the factorization scale, perturbative order, LCDA input, and power-correction
model before it is a QCD statement.

\begin{controlledapproximation}[Brodsky--Farrar scaling]
\label{ca:qcd-brodsky-farrar-scaling-status}
Consider a fixed-angle exclusive process
\[
  H_1+H_2\longrightarrow H_3+H_4
\]
whose leading Fock components contain \(n_i\) collinear elementary fields in
the corresponding leading-twist light-ray operator, and set
\(n=\sum_i n_i\).  If a leading-power exclusive factorization theorem holds,
if endpoint configurations are power suppressed in the specified scheme, and
if the hard coefficient is evaluated at momentum transfer of order
\(\sqrt s\), then
\begin{equation}
  \mathcal M(s,t)\sim s^{\,2-n/2}
  \quad\text{at fixed }t/s,
  \qquad
  \frac{\dd\sigma}{\dd t}\sim s^{\,2-n}.
  \label{eq:qcd-brodsky-farrar-counting}
\end{equation}
The derivation is the power counting of the hard coefficient after all
external hadrons have been replaced by their normalized leading-twist
distribution amplitudes.  It is not a dimensional-analysis argument applied to
hadron fields.  For example, a pion electromagnetic form factor has
\(n=4\) across the two pion light-ray operators and therefore scales as
\(F_\pi(Q^2)\sim Q^{-2}\), while fixed-angle
\(\pi\pi\to\pi\pi\) has \(n=8\) and gives
\(\dd\sigma/\dd t\sim s^{-6}\), up to logarithmic evolution and controlled
power corrections.
\end{controlledapproximation}

The pion \(S\)-matrix is thus accessed by different controlled descriptions in
different regions.  Chapter~\ref{ch:volume-ii-21} derives the chiral
low-energy amplitudes and anomaly couplings; Chapter~\ref{ch:volume-ii-07}
develops the analytic and dispersive language needed for the resonance
region; the present section gives the asymptotic exclusive regime governed by
light-ray factorization and asymptotic freedom.  These are not three theories.
They are three coordinate systems on matrix elements of the same QCD
operators, with different hypotheses controlling their errors.  A complete
nonperturbative construction of QCD would define the relevant matrix elements
first; chiral perturbation theory, dispersive methods, lattice calculations,
and hard exclusive factorization are then approximation or reconstruction
methods applied to that common object.

\section{Local Twist-Two Operators and Integer Moments}

The light-ray definition packages an infinite family of local operators.  For
quarks, the symmetric traceless spin-\(N\) operators have the form
\[
  \mathcal O_{q}^{\mu_1\cdots\mu_N}
  =
  \bar q\,
  \gamma^{(\mu_1}
  \ii D^{\mu_2}\cdots \ii D^{\mu_N)}
  q
  -
  \text{traces},
  \qquad N\ge1,
\]
with gauge-covariant derivatives in the appropriate representation.  The
parentheses denote symmetrization over Lorentz indices.  The gluon operators
are
\[
  \mathcal O_{g}^{\mu_1\cdots\mu_N}
  =
  F^{(\mu_1}{}_{\rho}\,
  \ii D^{\mu_2}\cdots\ii D^{\mu_{N-1}}
  F^{\mu_N)\rho}
  -
  \text{traces},
  \qquad N\ge2 .
\]
Both have twist two after subtracting traces.  Their hadron matrix elements
are fixed by Lorentz covariance up to scalar reduced matrix elements:
\[
  \bra{P}\mathcal O_{a}^{\mu_1\cdots\mu_N}(\mu)\ket{P}
  =
  2\,a_{a,N}(\mu)
  \bigl(P^{\mu_1}\cdots P^{\mu_N}-\text{traces}\bigr),
  \qquad a=q,g .
\]
The normalization of \(a_{a,N}\) is fixed already at tree level by replacing
the hadron with a single parton state.  Let
\(\ket{p,i,s}\) be a massless quark of flavor \(i\) and spin \(s\), and use
\(\bar u_s(p)\gamma^\mu u_s(p)=2p^\mu\).  Since each covariant derivative in
\(\mathcal O_{q_j}\) acts on the external plane wave as multiplication by
\(p^\mu\) at tree level,
\[
  \bra{p,i,s}\mathcal O_{q_j}^{\mu_1\cdots\mu_N}\ket{p,i,s}_{\rm tree}
  =
  2\delta_{ij}
  \bigl(p^{\mu_1}\cdots p^{\mu_N}-\text{traces}\bigr).
\]
Thus a quark target has
\[
  a_{q_j,N}^{\,q_i,{\rm tree}}=\delta_{ij},
  \qquad
  a_{g,N}^{\,q_i,{\rm tree}}=0 .
\]
For a single gluon state \(\ket{p,A,\lambda}\), the linearized field strength
is
\[
  F^{\mu\nu,A}(x)
  =
  \ii\bigl(p^\mu\varepsilon^\nu_\lambda
  -p^\nu\varepsilon^\mu_\lambda\bigr)\ee^{-\ii p\cdot x},
  \qquad
  p\cdot\varepsilon_\lambda=0 .
\]
The color- and spin-averaged partonic matrix element is
\[
  \frac{1}{2(N_c^2-1)}
  \sum_{A,\lambda}
  \bra{p,A,\lambda}\mathcal O_{g}^{\mu_1\cdots\mu_N}
  \ket{p,A,\lambda}_{\rm tree}
  =
  2
  \bigl(p^{\mu_1}\cdots p^{\mu_N}-\text{traces}\bigr),
\]
so
\[
  a_{g,N}^{\,g,{\rm tree}}=1,
  \qquad
  a_{q_j,N}^{\,g,{\rm tree}}=0 .
\]
Equivalently, the tree-level parton distributions are
\(f_{q_j/q_i}(x)=\delta_{ij}\delta(1-x)\),
\(f_{g/g}(x)=\delta(1-x)\), and the mixed quark-gluon distributions vanish.

Expanding the light-ray operator near zero separation gives these local
operators contracted with the lightlike vector \(n^\mu\).  Consequently the
integer moments of the parton distributions are the same reduced matrix
elements, with the precise flavor and charge-conjugation combination fixed by
the current channel:
\[
  a_{a,N}(\mu)
  =
  \int_0^1 \dd x\,x^{N-1} f_{a/H}(x,\mu)
\]
for the corresponding parton combination.  The local OPE and the nonlocal
light-ray factorization are therefore two presentations of the same
short-distance data.

Renormalization mixes operators with the same spin, twist, and quantum
numbers.  Taking integer moments of \eqref{eq:dglap-evolution} gives
\[
  \mu\frac{\dd}{\dd\mu}a_{a,N}(\mu)
  =
  \sum_b P_{ab,N}(g(\mu))\,a_{b,N}(\mu),
  \qquad
  P_{ab,N}(g)=\int_0^1\dd x\,x^{N-1}P_{ab}(x,g).
\]
The coefficient functions in the current OPE obey the dual evolution so that
the physical structure functions are independent of the auxiliary
factorization scale.  This is the local-operator form of perturbative
scaling violation in deep inelastic scattering.

For a spin-\(N\) leading-twist operator whose scaling dimension near the
ultraviolet is
\[
  \Delta_N(g)=N+2+\tau_{1,N}g^2+O(g^4),
\]
the one-loop running
\[
  \frac{\dd}{\dd\log\mu}\frac1{g^2(\mu)}
  =
  \frac{b}{2\pi^2}+O(g^2)
\]
implies logarithmic scaling of its matrix element and Wilson coefficient.
Equivalently, with
\[
  \omega:=\frac{2P\cdot q}{Q^2}=\frac1{x_B},
\]
the leading large-\(Q\) term in the \(N\)-th moment contribution has the form
\begin{equation}
  W_N(Q,\omega)
  \propto
  \omega^N
  \left(
    \log\frac{Q}{\Lambda}
  \right)^{-2\pi^2\tau_{1,N}/b}
  \left[1+O\!\left(\frac1{\log(Q/\Lambda)}\right)\right],
  \label{eq:qcd-dis-logarithmic-scaling-violation}
\end{equation}
up to the matrix of anomalous dimensions when several operators mix.  The
power \(\omega^N\) is the Bjorken-scaling kinematic dependence of a spin-\(N\)
operator; asymptotic freedom turns exact scaling into logarithmic scaling
violation.

\section{Cusped Wilson Lines, Large Spin, and the DIS Endpoint}
\label{sec:qcd-cusped-wilson-lines-large-spin}

The compact-\(x\) factorization statement in
Hypothesis~\ref{hyp:qcd-dis-leading-twist-factorization} deliberately does not
claim uniform control as \(x_B\to1\).  In that endpoint region the invariant
mass of the hadronic final state,
\[
  W_X^2=(P+q)^2
  =
  M^2+Q^2\left(\frac1{x_B}-1\right),
\]
is much smaller than \(Q^2\).  The short-distance coefficient functions then
contain distributions supported near \(x=1\), such as
\([\log^k(1-x)/(1-x)]_+\), and the perturbative expansion must be reorganized.
The universal coefficient controlling the leading endpoint singularity is
most cleanly defined by a cusped Wilson-line operator.  This definition is an
operator statement, not a parton-model picture.

Let \(R\) be a finite-dimensional unitary representation of \(SU(N_c)\), with
quadratic Casimir \(C_R\) in the trace-delta convention
\eqref{eq:qcd-casimir-definitions}.  In Euclidean signature choose two unit
vectors \(e_1,e_2\) with \(e_1\cdot e_2=\cos\phi\), \(0<\phi<\pi\), and let
\(\Gamma_{\phi,L,\epsilon}\) be the two-segment path
\[
  x(s)=-s e_1,\quad \epsilon\le s\le L,
  \qquad
  x(t)=t e_2,\quad \epsilon\le t\le L,
\]
joined at the origin after removing a ball of radius \(\epsilon\).  The
regulated cusped line is the path-ordered transporter
\[
  W_R(\Gamma_{\phi,L,\epsilon})
  =
  \mathcal P
  \exp\left\{\ii\int_{\Gamma_{\phi,L,\epsilon}}
  A_\mu^a(x)t_R^a\,\dd x^\mu\right\}.
\]
The straight-line endpoint and perimeter renormalizations are fixed by the
same regulator applied to a smooth line.  After these local smooth-line
renormalizations are divided out, the remaining logarithmic singularity at
the corner defines the Euclidean cusp coefficient \(K_R(\phi,g)\) by
\begin{equation}
  -\log\langle W_R(\Gamma_{\phi,L,\epsilon})\rangle_{\rm smooth-subtracted}
  =
  K_R(\phi,g)\log\frac{L}{\epsilon}+O(1).
  \label{eq:qcd-euclidean-cusp-coefficient-definition}
\end{equation}
The minus sign in this definition is part of the convention: it combines with
the \(i^2\) from expanding the Wilson transporter, so that the one-loop
coefficient below is read directly from the positive Euclidean propagator.
For timelike Wilson lines in Lorentzian signature we instead use the
positive lightlike cusp anomalous dimension
\[
  \Gamma_{\rm cusp}^R(g)
  :=
  \lim_{\chi\to\infty}
  \frac{\Gamma_{\rm cusp}^R(\chi,g)}{\chi},
  \qquad
  -v_1\cdot v_2=\cosh\chi,\quad v_1^2=v_2^2=-1,
\]
where \(\Gamma_{\rm cusp}^R(\chi,g)\) is obtained from the same local
renormalization problem by the analytic continuation \(\phi=\ii\chi\).
This convention is the one used in
\eqref{eq:qcd-quark-cusp-one-loop-trace-delta}.

At one loop the angular dependence follows from a single eikonal exchange.
In Feynman gauge, with the connection normalization of this chapter, the
Euclidean short-distance propagator is
\[
  \langle A_\mu^a(x)A_\nu^b(0)\rangle
  =
  \frac{g^2\delta^{ab}\delta_{\mu\nu}}{4\pi^2x^2}+O(g^4).
\]
The exchange between the two rays contributes, up to cutoff-independent
finite terms, the following scale-invariant part of
\(-\log\langle W_R\rangle\):
\[
  \frac{g^2C_R}{4\pi^2}
  \int_\epsilon^L\dd s\int_\epsilon^L\dd t\,
  \frac{e_1\cdot e_2}
  {s^2+t^2+2st\,e_1\cdot e_2}.
\]
The coefficient of \(\log(L/\epsilon)\) is local at the cusp, so set
\(t=sr\) in the scale-invariant part.  The angular integral is
\begin{equation}
  J(\phi)
  :=
  \int_0^\infty
  \frac{\cos\phi\,\dd r}
  {r^2+2r\cos\phi+1}
  =
  \phi\cot\phi .
  \label{eq:qcd-cusp-angular-integral}
\end{equation}
Indeed,
\[
  r^2+2r\cos\phi+1
  =
  (r+\cos\phi)^2+\sin^2\phi,
\]
and the antiderivative is
\[
  \cot\phi\,
  \arctan\frac{r+\cos\phi}{\sin\phi}.
\]
Taking \(r\) from \(0\) to \(\infty\) gives
\(\cot\phi(\pi/2-\arctan(\cot\phi))=\phi\cot\phi\), with the principal branch
of \(\arctan\) for \(0<\phi<\pi\).  A smooth straight line corresponds to
\(\phi=0\), for which \(J(0)=1\).  Therefore the corner-specific logarithm is
\begin{equation}
  K_R^{(1)}(\phi,g)
  =
  \frac{g^2C_R}{4\pi^2}
  \bigl(\phi\cot\phi-1\bigr).
  \label{eq:qcd-euclidean-cusp-one-loop}
\end{equation}
Continuing \(\phi=\ii\chi\) gives the Lorentzian timelike cusp coefficient
\begin{equation}
  \Gamma_{\rm cusp}^R(\chi,g)
  =
  \frac{g^2C_R}{4\pi^2}
  \bigl(\chi\coth\chi-1\bigr)+O(g^4),
  \qquad
  \Gamma_{\rm cusp}^R(g)
  =
  \frac{g^2C_R}{4\pi^2}+O(g^4).
  \label{eq:qcd-lorentzian-cusp-one-loop}
\end{equation}
For \(R=\square\), this is exactly the quark coefficient in
\eqref{eq:qcd-quark-cusp-one-loop-trace-delta}.

The same coefficient appears in the endpoint renormalization of light-ray
operators.  Near \(x=1\), the emitted radiation is soft compared with the
parent quark momentum, and the Wilson transporter in the light-ray operator
is approximated by eikonal propagation along the two nearly lightlike
directions selected by the incoming hadron and the hard current.  The local
renormalization of the resulting lightlike cusp supplies the coefficient of
the plus distribution \(D_0=(1-x)^{-1}_+\) in the nonsinglet kernel:
\begin{equation}
  P_{\rm ns}(x,g)
  =
  \Gamma_{\rm cusp}^q(g)\,D_0(x)
  +\text{terms with less singular Mellin growth as }x\to1 .
  \label{eq:qcd-endpoint-kernel-cusp-coefficient}
\end{equation}
Equation \eqref{eq:qcd-endpoint-kernel-cusp-coefficient} is meant as a
perturbative renormalization statement for the specified light-ray operator
and subtraction scheme.  It is not a nonperturbative theorem about the full
endpoint region of every DIS observable; threshold factorization also
requires a hard coefficient, a jet function, a soft matrix element, a
rapidity prescription when appropriate, and a controlled power remainder.

The Mellin transform turns this endpoint problem into a large-spin problem.
For a distribution \(K(x)\), define
\[
  K_N:=\int_0^1\dd x\,x^{N-1}K(x).
\]
Using the \(D_0\) convention above, the nonsinglet one-loop kernel has the
exact moment
\begin{equation}
  P_{{\rm ns},N}^{(0)}
  =
  C_F\left[
    -2H_{N-1}-\frac1N-\frac1{N+1}+\frac32
  \right],
  \qquad N\ge1 .
  \label{eq:qcd-nonsinglet-moment-exact}
\end{equation}
The harmonic number appears because
\begin{equation}
  \int_0^1\dd x\,x^{N-1}D_0(x)
  =
  \int_0^1\dd x\,
  \frac{x^{N-1}-1}{1-x}
  =
  -H_{N-1}.
  \label{eq:qcd-d0-mellin-harmonic}
\end{equation}
The last equality follows by writing
\((1-x^{N-1})/(1-x)=1+x+\cdots+x^{N-2}\) before integration.  Thus
\eqref{eq:qcd-nonsinglet-moment-exact} gives quark-number conservation at
\(N=1\), and at large \(N\)
\begin{equation}
  P_{{\rm ns},N}^{(0)}
  =
  -2C_F\log N
  +C_F\left(\frac32-2\gamma_{\rm E}\right)
  +O(N^{-1}).
  \label{eq:qcd-nonsinglet-large-n}
\end{equation}
Therefore, with \(P=(g^2/8\pi^2)P^{(0)}+O(g^4)\),
\begin{equation}
  P_{{\rm ns},N}(g)
  =
  -\Gamma_{\rm cusp}^q(g)\log N+O(N^0),
  \qquad
  \Gamma_{\rm cusp}^q(g)
  =
  \frac{g^2C_F}{4\pi^2}+O(g^4).
  \label{eq:qcd-large-spin-dglap-cusp}
\end{equation}
This is the evolution eigenvalue for matrix elements in the convention
\(\mu\,\dd a_N/\dd\mu=P_Na_N\).  If instead one writes the anomalous dimension
of the local twist-two operator itself by
\(\mu\,\dd\mathcal O_N/\dd\mu=-\gamma_N^{\mathcal O}\mathcal O_N+\cdots\),
then
\[
  \gamma_N^{\mathcal O}(g)
  =
  \Gamma_{\rm cusp}^q(g)\log N+O(N^0).
\]
The sign difference is bookkeeping: coefficient functions and operator
matrix elements obey dual renormalization-group equations so that the
physical hadronic tensor is independent of the factorization scale.

The product \(g^2C_R\) in
\eqref{eq:qcd-lorentzian-cusp-one-loop} is invariant under generator-coordinate
changes.  If \(T^a\) denotes the common half-trace generator and
\(g_{\rm ht}=\sqrt2\,g\), then
\[
  g^2C_F=g_{\rm ht}^2C_F^{\rm ht}.
\]
The cusp coefficient is thus independent of the choice of generator
coordinates once the coupling coordinate is transformed with the action.

Endpoint resummation is not a new definition of the PDF.  It is an additional
factorization of the coefficient function in a region where the compact-\(x\)
remainder estimate is not uniform.  A precise threshold factorization theorem
must specify the hard current coefficient, the jet function for the nearly
collinear final-state radiation, the soft Wilson-line matrix element, the
contour prescription for Mellin or Laplace inversion, and a remainder estimate
in the joint limit \(Q\to\infty\), \(N\to\infty\).  The large-spin logarithm in
\eqref{eq:qcd-large-spin-dglap-cusp} is the operator-renormalization side of
that structure.  The same Wilson-line cusp observable is the QCD analogue of
the large-spin cusp scaling function studied in
Chapter~\ref{ch:planar-n4-asymptotic-bethe-ansatz}, but the logical status is
different: in QCD it is a perturbative Wilson-line anomalous dimension tied
to factorization hypotheses, while in planar \(\mathcal N=4\) super
Yang--Mills the BES equation gives an integrability construction of the
planar large-spin function in a special conformal theory.  The small-\(x\)
limit is a different boundary of the same DIS observable and is not obtained
by continuing the threshold expansion.

\begin{controlledapproximation}[Tested DIS threshold plus-distribution window]
\label{ca:qcd-dis-threshold-plus-distribution-window}
Let
\[
  D_k(x):=\left[\frac{\log^k(1-x)}{1-x}\right]_+,\qquad k\ge0,
\]
so that, for a smooth test function \(\varphi\),
\begin{equation}
  \langle D_k,\varphi\rangle
  =
  \int_0^1
  \frac{\log^k(1-x)}{1-x}
  \bigl(\varphi(x)-\varphi(1)\bigr)\,\dd x .
  \label{eq:qcd-dis-threshold-plus-action}
\end{equation}
The subtraction is the physical endpoint datum.  The cusp kernel sees the
variation of the measured weight near \(x=1\), not the value of the inclusive
bin at the elastic endpoint.  On a terminal cell \(1-\delta\le x\le1\), with
\(0<\delta<1\),
\begin{align}
  \left|\int_{1-\delta}^1
  \frac{\varphi(x)-\varphi(1)}{1-x}\,\dd x\right|
  &\le
  \|\varphi'\|_{L^\infty([1-\delta,1])}\,\delta,
  \label{eq:qcd-dis-d0-terminal-bound}\\
  \left|\int_{1-\delta}^1
  \frac{\log(1-x)}{1-x}
  \bigl(\varphi(x)-\varphi(1)\bigr)\,\dd x\right|
  &\le
  \|\varphi'\|_{L^\infty([1-\delta,1])}\,
  \delta\bigl(1+|\log\delta|\bigr).
  \label{eq:qcd-dis-d1-terminal-bound}
\end{align}
Equivalently, writing \(u=1-x\) and
\(\varphi(1-u)=\varphi(1)-a_1u+a_2u^2+O(u^3)\), the terminal cell contributes
\begin{align}
  \langle D_0,\varphi\rangle_{\delta}
  &=
  -a_1\delta+\frac{a_2}{2}\delta^2+O(\delta^3),
  \label{eq:qcd-dis-d0-terminal-taylor}\\
  \langle D_1,\varphi\rangle_{\delta}
  &=
  -a_1\delta(\log\delta-1)
  +\frac{a_2}{2}\delta^2\left(\log\delta-\frac12\right)
  +O(\delta^3|\log\delta|).
  \label{eq:qcd-dis-d1-terminal-taylor}
\end{align}
Thus the Mellin identity
\[
  \langle D_0,x^{N-1}\rangle=-H_{N-1}
  =
  -\log N-\gamma_{\rm E}+O(N^{-1})
\]
is a large-\(N\) probe of a moving family of test functions.  It is not a
uniform consequence of the compact-\(x\) factorization statement on fixed test
functions supported away from \(x=1\).

A threshold theorem for DIS must therefore carry more data than the cusp
coefficient.  A moment-space coordinate for the approximation datum is
\begin{equation}
  F_{i,N}^{\rm thr}(Q)
  =
  H_i(Q,\mu_h)\,
  U_H\,
  J_{i,N}(Q,\mu_j)\,
  U_J\,
  S_{i,N}(\mu_s)\,
  U_S\,
  f_N(\mu_f)
  +R_{i,N}^{\rm thr},
  \label{eq:qcd-dis-threshold-factorized-coordinate}
\end{equation}
where the hard current, final-state jet, soft Wilson-line matrix element,
evolution kernels, Mellin-inversion contour or prescription, and remainder
\(R_{i,N}^{\rm thr}\) are part of the approximation datum.  The cusp anomalous
dimension controls the leading double-logarithmic evolution shared by these
factors; it does not by itself determine the finite matching constants, the
soft measurement convention, the prescription near the Landau boundary, or the
power remainders such as \(1/N\) and \(N\Lambda_{\rm QCD}/Q\).
\end{controlledapproximation}

\section{\texorpdfstring{Small-\(x\) Boundary, Dipoles, and the BFKL Kernel}
{Small-x Boundary, Dipoles, and the BFKL Kernel}}
\label{sec:qcd-small-x-bfkl}

The compact-\(x_B\) factorization hypothesis in
Hypothesis~\ref{hyp:qcd-dis-leading-twist-factorization} also does not claim
uniform control as \(x_B\to0\).  Put
\[
  Y:=\log\frac1{x_B},
\]
up to a process-dependent finite shift fixed by the chosen rapidity
factorization scheme.  In the high-energy boundary \(Y\gg1\), powers of
\(g^2Y\) may compete with powers of \(g^2\).  The leading-logarithmic
small-\(x\) expansion is therefore not a new definition of QCD and not a
replacement for the hadronic tensor
\eqref{eq:hadronic-tensor-definition}.  It is a rapidity factorization of
regulated Wilson-line matrix elements in a kinematic region where the
ordinary twist-two evolution at fixed \(x_B\) is not uniform.
The primary sources for the leading high-energy kernel and its Mellin
eigenvalue are the BFKL papers of Kuraev--Lipatov--Fadin and
Balitsky--Lipatov
\cite{KuraevFadinLipatov1977,BalitskyLipatov1978}.

Let \(\mathbf x,\mathbf y,\mathbf z\in\mathbb R^2\) denote transverse
coordinates, and set
\[
  r_{xy}:=|\mathbf x-\mathbf y|.
\]
The leading transverse kernel is
\begin{equation}
  M_{\mathbf x\mathbf y\mathbf z}
  :=
  \frac{r_{xy}^2}{r_{xz}^2 r_{zy}^2}.
  \label{eq:qcd-dipole-transverse-kernel}
\end{equation}
The singularities at \(\mathbf z=\mathbf x\) and
\(\mathbf z=\mathbf y\) are not meaningful as separate real-emission
probabilities.  They are defined only after the virtual subtraction in the
evolution equation, or after an equivalent rapidity regulator and finite-part
map have been specified.

\begin{definition}[Regulated small-\(x\) dipole setup]
\label{def:qcd-small-x-dipole-setup}
A small-\(x\) dipole setup for a color-singlet probe consists of:
\begin{enumerate}
  \item a QCD regulator \(\mathcal R_\epsilon\) preserving background gauge
  covariance, or a gauge-fixed BV regulator with the quantum master equation;
  \item two nearly lightlike Wilson-line directions, a rapidity regulator
  \(\rho\), and a finite rapidity separation \(Y\);
  \item fundamental Wilson transporters
  \(U_{\rho,Y}(\mathbf x)\) and \(U_{\rho,Y}(\mathbf y)\), with transverse
  closures or boundary conditions that make the residual endpoint gauge
  transformations explicit;
  \item the renormalized color dipole matrix element
  \begin{equation}
    S_Y(\mathbf x,\mathbf y)
    :=
    \frac1{N_c}
    \left\langle
      \operatorname{tr}_{\square}
      \bigl[
        U_{\rho,Y}(\mathbf x)U_{\rho,Y}(\mathbf y)^\dagger
      \bigr]
    \right\rangle_{\epsilon,\rho,Y}^{\rm ren},
    \qquad
    N_Y(\mathbf x,\mathbf y):=1-S_Y(\mathbf x,\mathbf y);
    \label{eq:qcd-small-x-dipole-definition}
  \end{equation}
  \item a finite-part prescription for ultraviolet endpoint divergences,
  lightlike or near-lightlike rapidity divergences, and possible cusp
  divergences introduced by the chosen closures.
\end{enumerate}
The quantity \(S_Y\) is a Wilson-line matrix element.  Its connection to a
measured small-\(x\) structure function is an additional high-energy
factorization statement with its own impact factor, rapidity subtraction, and
power-remainder data.
\end{definition}

At finite regulator, the trace in
\eqref{eq:qcd-small-x-dipole-definition} is gauge invariant after the
endpoint closures or boundary conditions of
Definition~\ref{def:qcd-small-x-dipole-setup} are included.  A background
gauge-covariant finite-part prescription therefore gives a gauge-invariant
renormalized dipole matrix element.
Each Wilson transporter transforms by multiplication with the gauge
transformation at its two endpoints.  Along a closed contour the adjacent
endpoint factors cancel and the residual base-point factor is removed by the
cyclicity of \(\operatorname{tr}_{\square}\).  If the regulated Wilson lines
are represented instead as open lines ending at a rapidity boundary, the
boundary condition is part of the setup and supplies the same endpoint
cancellation.  The finite-part map changes only regulator and scale
coordinates and is assumed background gauge covariant; hence it cannot
introduce an uncanceled color index.

\begin{hypothesis}[Leading-logarithmic dipole/BFKL evolution setup]
\label{hyp:qcd-leading-log-small-x-setup}
In a fixed-coupling leading-logarithmic rapidity scheme, the linearized
color-singlet dipole amplitude obeys
\begin{equation}
  \partial_Y N_Y(\mathbf x,\mathbf y)
  =
  \frac{g^2 C_A}{8\pi^3}
  \int_{\mathbb R^2}\dd^2\mathbf z\,
  M_{\mathbf x\mathbf y\mathbf z}
  \bigl[
    N_Y(\mathbf x,\mathbf z)
    +N_Y(\mathbf z,\mathbf y)
    -N_Y(\mathbf x,\mathbf y)
  \bigr],
  \label{eq:qcd-coordinate-bfkl-equation}
\end{equation}
where \(C_A=2N_c\) in the trace convention of
\eqref{eq:qcd-casimir-definitions}.  Equivalently, with
\[
  \bar\alpha_{\rm s}:=\frac{g^2C_A}{4\pi^2}
  =
  \frac{g^2N_c}{2\pi^2},
\]
the right-hand side is
\[
  \bar\alpha_{\rm s}\,
  \frac1{2\pi}
  \int\dd^2\mathbf z\,M_{\mathbf x\mathbf y\mathbf z}[\cdots].
\]
In the common half-trace convention,
\(g_{\rm ht}^2=2g^2\) and \(C_A^{\rm ht}=N_c\), so the product
\(g^2C_A=g_{\rm ht}^2C_A^{\rm ht}\) and the kernel coefficient are unchanged.

For a large-\(N_c\) dipole-factorization truncation of the Balitsky hierarchy,
the corresponding nonlinear equation for \(N_Y=1-S_Y\) has the additional
term
\begin{equation}
  -\frac{g^2 C_A}{8\pi^3}
  \int\dd^2\mathbf z\,
  M_{\mathbf x\mathbf y\mathbf z}\,
  N_Y(\mathbf x,\mathbf z)N_Y(\mathbf z,\mathbf y).
  \label{eq:qcd-bk-nonlinear-term}
\end{equation}
At finite \(N_c\), the exact Wilson-line rapidity hierarchy contains higher
multi-trace operators; replacing them by the product in
\eqref{eq:qcd-bk-nonlinear-term} is an additional closure assumption, not a
theorem about continuum QCD.
The dipole picture and its large-\(N_c\) small-\(x\) DIS evolution lineage are
traced through Mueller's onium construction, Balitsky's Wilson-line operator
expansion, and Kovchegov's \(F_2\) evolution equation
\cite{Mueller1994SoftGluons,Balitsky1996HighEnergy,
Kovchegov1999SmallXF2}.
\end{hypothesis}

\subsection{Finite Wilson-Line Evolution and the JIMWLK Theorem Boundary}
\label{subsec:qcd-finite-jimwlk-boundary}

The preceding dipole equation is a closed equation only after a truncation has
been chosen.  The more invariant small-\(x\) object is the rapidity evolution
of the whole algebra generated by Wilson lines.  To state the mathematical
content without importing the usual folklore terminology, first put the
Wilson-line variables on a finite transverse regulator.
The target-side Wilson renormalization group and its functional
Fokker--Planck form are the JIMWLK development
\cite{JalilianMarianKovnerLeonidovWeigert1998HighDensity,
JalilianMarianKovnerWeigert1998FiniteDensity,
IancuLeonidovMcLerran2001CGC,Weigert2002Unitarity}.

\begin{definition}[Finite Wilson-line rapidity-evolution generator]
\label{def:qcd-finite-wilson-line-rapidity-datum}
Let \(\Lambda_\perp\) be a finite set of transverse sites and let
\begin{equation}
  \mathcal C_{\Lambda_\perp}:=SU(N_c)^{\Lambda_\perp}
\end{equation}
with product Haar measure \(\dd h(U)\).  For
\(\mathbf x\in\Lambda_\perp\) and Lie-algebra index \(a\), let
\(L_{\mathbf x}^a\) be the left-invariant vector field
\begin{equation}
  (L_{\mathbf x}^aF)(U)
  :=
  \left.
  \frac{\dd}{\dd t}
  F(\ldots,\ee^{\ii t t^a}U_{\mathbf x},\ldots)
  \right|_{t=0},
  \label{eq:qcd-finite-wilson-left-derivative}
\end{equation}
where the generators obey the trace convention of
\eqref{eq:qcd-casimir-definitions}.  A finite Wilson-line
rapidity-evolution structure consists of a family of smooth coefficients
\[
  \mathcal A_{\mathbf x\mathbf y}^{ab}(U;Y)
  =
  \mathcal A_{\mathbf y\mathbf x}^{ba}(U;Y)
\]
whose finite matrix
\((\mathcal A_{\mathbf x\mathbf y}^{ab})\) is nonnegative at each
\((U,Y)\), is equivariant under residual global color conjugation, and is
defined in the same rapidity and ultraviolet regulator scheme as the Wilson
lines.  The corresponding finite Wilson-line generator is the
second-order operator
\begin{equation}
  (\mathcal H_YF)(U)
  :=
  \frac12
  \sum_{\mathbf x,\mathbf y\in\Lambda_\perp}
  \sum_{a,b}
  L_{\mathbf x}^a
  \left[
    \mathcal A_{\mathbf x\mathbf y}^{ab}(U;Y)
    L_{\mathbf y}^bF(U)
  \right].
  \label{eq:qcd-finite-wilson-line-generator}
\end{equation}
This finite generator is a finite-regulator replacement for the JIMWLK Hamiltonian only
after the coefficient \(\mathcal A\) has been derived from a specified QCD
rapidity factorization scheme.
\end{definition}

The finite object has two equivalent presentations.  In the weak, or
observable, presentation a rapidity-dependent state \(W_Y\dd h\) on
\(\mathcal C_{\Lambda_\perp}\) obeys
\begin{equation}
  \frac{\dd}{\dd Y}
  \int_{\mathcal C_{\Lambda_\perp}}F(U)\,W_Y(U)\,\dd h(U)
  =
  \int_{\mathcal C_{\Lambda_\perp}}(\mathcal H_YF)(U)\,W_Y(U)\,\dd h(U)
  \label{eq:qcd-finite-jimwlk-weak-form}
\end{equation}
for every smooth gauge-invariant Wilson-line observable \(F\).  In the strong,
or density, presentation
\begin{equation}
  \partial_YW_Y=\mathcal H_Y^*W_Y,
  \qquad
  \mathcal H_Y^*w
  =
  \frac12
  \sum_{\mathbf x,\mathbf y,a,b}
  L_{\mathbf y}^b
  \left[
    \mathcal A_{\mathbf x\mathbf y}^{ab}(U;Y)
    L_{\mathbf x}^aw
  \right],
  \label{eq:qcd-finite-jimwlk-density-form}
\end{equation}
where the adjoint is taken with respect to product Haar measure.  The two
forms are identical by integration by parts on the compact Lie group:
\[
  \int (L_{\mathbf x}^aF)G\,\dd h
  =
  -\int F(L_{\mathbf x}^aG)\,\dd h .
\]
Applying this identity twice to
\eqref{eq:qcd-finite-wilson-line-generator} gives
\eqref{eq:qcd-finite-jimwlk-density-form}.  Since
\(\mathcal H_Y1=0\), the total mass \(\int W_Y\dd h\) is conserved.  Since
the coefficient matrix is nonnegative, the principal symbol is that of a
diffusion on the compact manifold.  When the closure of \(\mathcal H_Y\) on
continuous functions satisfies the positive maximum principle, the
Hille--Yosida theorem gives the corresponding Markov semigroup.  This
semigroup assertion belongs to ordinary finite-dimensional diffusion analysis;
the QCD-specific work is the derivation and limiting control of
\(\mathcal A\).  These are finite statements about
\(\mathcal C_{\Lambda_\perp}\), not continuum QCD theorems.

For the dipole observable
\begin{equation}
  F_{\mathbf x\mathbf y}(U)
  :=
  \frac1{N_c}
  \operatorname{tr}_{\square}(U_{\mathbf x}U_{\mathbf y}^\dagger),
  \label{eq:qcd-finite-dipole-observable}
\end{equation}
the weak equation
\eqref{eq:qcd-finite-jimwlk-weak-form} generates the first equation of the
Balitsky hierarchy.  The right-hand side contains Wilson-line observables with
more transverse labels.  A closed BK equation is obtained only after
large-\(N_c\), mean-field, or other factorization input has replaced those
higher observables by products of dipoles.  Thus the logical chain is
\[
  \text{finite Wilson-line diffusion}
  \longrightarrow
  \text{Balitsky hierarchy}
  \longrightarrow
  \text{BK closure under extra hypotheses},
\]
not a direct theorem from the BFKL eigenvalue calculation.

\paragraph{Dipole projection and the BK closure boundary.}
The first arrow in the preceding display is an operator statement on the
finite Wilson-line algebra; the second arrow is a closure operation on a
subsystem of expectation values.  It is useful to write this subsystem before
passing to the continuum transverse plane.  Fix a finite transverse set
\(\Lambda_\perp\) and nonnegative cell weights
\(\omega_{\mathbf x\mathbf y;\mathbf z}(Y)\), symmetric under
\(\mathbf x\leftrightarrow\mathbf y\), which represent the regulated
cell-integrals of the leading dipole kernel in the same rapidity scheme as
\(\mathcal H_Y\).  For distinct transverse labels set
\[
  S_{\mathbf x\mathbf y}^{(1)}(Y)
  :=
  \langle F_{\mathbf x\mathbf y}\rangle_Y,
  \qquad
  S_{\mathbf x\mathbf z;\mathbf z\mathbf y}^{(2)}(Y)
  :=
  \langle
    F_{\mathbf x\mathbf z}F_{\mathbf z\mathbf y}
  \rangle_Y ,
\]
where the expectation is the weak Wilson-line state in
\eqref{eq:qcd-finite-jimwlk-weak-form}.  A finite leading-logarithmic dipole
projection is the assertion that
\begin{equation}
  \partial_Y S_{\mathbf x\mathbf y}^{(1)}
  =
  \sum_{\mathbf z\in\Lambda_\perp\setminus\{\mathbf x,\mathbf y\}}
  \omega_{\mathbf x\mathbf y;\mathbf z}
  \left(
    S_{\mathbf x\mathbf z;\mathbf z\mathbf y}^{(2)}
    -
    S_{\mathbf x\mathbf y}^{(1)}
  \right).
  \label{eq:qcd-finite-balitsky-first-dipole-projection}
\end{equation}
Equation~\eqref{eq:qcd-finite-balitsky-first-dipole-projection} is not a
closed equation for \(S^{(1)}\).  Its new variable is the double-dipole
expectation \(S^{(2)}\), whose rapidity derivative involves still higher
Wilson-line observables.  This is the finite-regulator form of the Balitsky
hierarchy.

The BK reduction is a controlled approximation only if one supplies a
decorrelation estimate for the connected double-dipole coordinate
\[
  C_{\mathbf x\mathbf z;\mathbf z\mathbf y}(Y)
  :=
  S_{\mathbf x\mathbf z;\mathbf z\mathbf y}^{(2)}(Y)
  -
  S_{\mathbf x\mathbf z}^{(1)}(Y)S_{\mathbf z\mathbf y}^{(1)}(Y).
\]
With
\begin{equation}
  E_{\mathbf x\mathbf y}^{\rm BK}(Y)
  :=
  \sum_{\mathbf z\ne\mathbf x,\mathbf y}
  \omega_{\mathbf x\mathbf y;\mathbf z}(Y)\,
  C_{\mathbf x\mathbf z;\mathbf z\mathbf y}(Y),
  \label{eq:qcd-bk-closure-error-coordinate}
\end{equation}
the exact first dipole equation is
\[
  \partial_Y S_{\mathbf x\mathbf y}^{(1)}
  =
  \sum_{\mathbf z\ne\mathbf x,\mathbf y}
  \omega_{\mathbf x\mathbf y;\mathbf z}
  \left(
    S_{\mathbf x\mathbf z}^{(1)}S_{\mathbf z\mathbf y}^{(1)}
    -
    S_{\mathbf x\mathbf y}^{(1)}
  \right)
  +E_{\mathbf x\mathbf y}^{\rm BK}.
\]
Thus the phrase ``large-\(N_c\) BK equation'' means that a separate argument
has made \(E^{\rm BK}\) negligible in the norm and rapidity interval under
discussion.  In the amplitude variable
\(N_{\mathbf x\mathbf y}^{(1)}=1-S_{\mathbf x\mathbf y}^{(1)}\), the closed
finite BK equation obtained by setting \(E^{\rm BK}=0\) is
\begin{equation}
  \partial_Y N_{\mathbf x\mathbf y}
  =
  \sum_{\mathbf z\ne\mathbf x,\mathbf y}
  \omega_{\mathbf x\mathbf y;\mathbf z}
  \left(
    N_{\mathbf x\mathbf z}
    +N_{\mathbf z\mathbf y}
    -N_{\mathbf x\mathbf y}
    -N_{\mathbf x\mathbf z}N_{\mathbf z\mathbf y}
  \right).
  \label{eq:qcd-finite-bk-amplitude-equation}
\end{equation}
This is the finite-cell version of
\eqref{eq:qcd-coordinate-bfkl-equation} plus
\eqref{eq:qcd-bk-nonlinear-term}.

The finite error estimate is elementary but important.  Suppose a real
mean-field solution \(\widetilde S_{\mathbf x\mathbf y}\) and the exact first
coordinates \(S_{\mathbf x\mathbf y}^{(1)}\) both take values in
\([0,1]\) on a rapidity interval \(0\le Y\le Y_*\), share the same initial
data, and have
\[
  L:=\sup_{Y\le Y_*}\sup_{\mathbf x\ne\mathbf y}
  \sum_{\mathbf z\ne\mathbf x,\mathbf y}
  \omega_{\mathbf x\mathbf y;\mathbf z}(Y)<\infty ,
  \qquad
  \varepsilon(Y):=
  \sup_{\mathbf x\ne\mathbf y}
  |E_{\mathbf x\mathbf y}^{\rm BK}(Y)|.
\]
Writing
\[
  d(Y):=
  \sup_{\mathbf x\ne\mathbf y}
  |S_{\mathbf x\mathbf y}^{(1)}(Y)-\widetilde S_{\mathbf x\mathbf y}(Y)|,
\]
one obtains
\begin{equation}
  d(Y)
  \le
  \int_0^Y \ee^{3L(Y-s)}\varepsilon(s)\,\dd s.
  \label{eq:qcd-finite-bk-closure-error-bound}
\end{equation}
Indeed, on the unit interval
\(|ab-a'b'|\le |a-a'|+|b-b'|\), so the product term contributes at most
\(2Ld(Y)\), the virtual term at most \(Ld(Y)\), and the connected covariance
term contributes \(\varepsilon(Y)\); Gronwall gives
\eqref{eq:qcd-finite-bk-closure-error-bound}.  The closed BK vector field
also points into the unit cube: at \(S_{\mathbf x\mathbf y}=0\) its right-hand
side is nonnegative, while at \(S_{\mathbf x\mathbf y}=1\) it is
nonpositive.  This finite ODE fact explains the black-disk and transparent
fixed points of the closed approximation.  It does not replace the missing
QCD estimate on \(E^{\rm BK}\), nor the continuum limit of the cell weights.

The continuum JIMWLK problem is the removal of the finite transverse regulator
and rapidity regulator from this finite compact-manifold construction while
keeping the Wilson-line observables, the coefficient kernel, and the
factorization map to the measured process in a topology strong enough to pass
to the limit in \eqref{eq:qcd-finite-jimwlk-weak-form}.  The monograph
therefore treats the color-glass weight functional as a renormalized
Wilson-line state, not as a literal probability measure on classical gluon
fields unless such a measure has been constructed in the chosen regulator.

\paragraph{Projective continuum limit for Wilson-line states.}
A continuum statement of JIMWLK must be stronger than a sequence of finite
Balitsky equations.  Let
\(\{\Lambda_\perp^{(m)}\}_{m\in I}\) be a directed family of finite transverse
regulators and let
\(\mathcal A_m=C^\infty(SU(N_c)^{\Lambda_\perp^{(m)}})^{G_{\rm diag}}\) be
the gauge-invariant Wilson-line algebra on the \(m\)-th regulator.  A
cylinder observable \(F\) depends on a fixed finite list of transverse labels;
for all sufficiently fine regulators it has a representative \(F_m\in
\mathcal A_m\).  A projective Wilson-line state limit on
\([0,Y_*]\) consists of
\[
  \bigl(
    \omega_{m,Y},\mathcal H_{m,Y},F\mapsto F_m,
    \omega_Y,\mathcal H^{\rm cyl}_Y,\mathcal I_Y
  \bigr)
\]
with the following data.  First, the finite states are compatible on cylinder
observables:
\begin{equation}
  \sup_{0\le Y\le Y_*}
  |\omega_{m,Y}(F_m)-\omega_Y(F)|
  \le \delta_{m,F},
  \qquad
  \delta_{m,F}\longrightarrow0 .
  \label{eq:qcd-jimwlk-projective-state-convergence}
\end{equation}
Second, the finite generators converge on the same cylinder test:
\begin{equation}
  \sup_{0\le Y\le Y_*}
  \|
    \mathcal H_{m,Y}F_m-(\mathcal H^{\rm cyl}_YF)_m
  \|_\infty
  \le \varepsilon_{m,F},
  \qquad
  \varepsilon_{m,F}\longrightarrow0 .
  \label{eq:qcd-jimwlk-projective-generator-convergence}
\end{equation}
Third, the rapidity dependence is uniformly controlled on cylinder tests, so
the finite weak equations are equicontinuous in \(Y\) and limits may be
passed through rapidity integrals.  Fourth, positivity, normalization, and the
positive-maximum-principle/Markov property hold at finite \(m\) and are closed
under the cylinder limit.  Fifth, \(\mathcal I_Y\) is the process map: it
turns the Wilson-line state into the measured small-\(x\) observable with the
chosen impact factors, subtractions, and power-remainder topology.

Under these hypotheses the finite weak equation has a genuine cylinder
limit.  Indeed, for \(Y_0<Y_1\),
\[
  \omega_{m,Y_1}(F_m)-\omega_{m,Y_0}(F_m)
  =
  \int_{Y_0}^{Y_1}
  \omega_{m,s}(\mathcal H_{m,s}F_m)\,\dd s .
\]
After adding and subtracting the limiting terms one obtains the error ledger
\begin{equation}
\begin{aligned}
&
\left|
  \omega_{Y_1}(F)-\omega_{Y_0}(F)
  -
  \int_{Y_0}^{Y_1}
    \omega_s(\mathcal H^{\rm cyl}_sF)\,\dd s
\right|
\\
&\qquad\le
  2\delta_{m,F}
  +
  \int_{Y_0}^{Y_1}
  \bigl(
    \delta_{m,\mathcal H^{\rm cyl}F}(s)
    +
    \varepsilon_{m,F}(s)
  \bigr)\,\dd s ,
\end{aligned}
\label{eq:qcd-jimwlk-projective-weak-limit-budget}
\end{equation}
where the right-hand side is written with the corresponding uniform or
integrable bound on the state error for the generator observable when the
errors depend on \(s\).  Letting \(m\to\infty\)
gives the continuum weak equation on the cylinder algebra,
\[
  \omega_{Y_1}(F)-\omega_{Y_0}(F)
  =
  \int_{Y_0}^{Y_1}
    \omega_s(\mathcal H^{\rm cyl}_sF)\,\dd s .
\]
If \(Y\mapsto\omega_Y(F)\) is absolutely continuous, this is the
differential weak JIMWLK equation for cylinder observables.  The BK closure
estimate above controls one projected dipole subsystem and one connected
covariance coordinate; it does not supply
\eqref{eq:qcd-jimwlk-projective-state-convergence},
\eqref{eq:qcd-jimwlk-projective-generator-convergence}, or the process map.
Thus continuum JIMWLK is a projective theorem about Wilson-line states and
QCD factorization; the formal functional Hamiltonian names only the candidate
cylinder generator.

\begin{controlledapproximation}[Measured small-\(x\) observable map]
\label{ca:qcd-small-x-measured-observable-map}
The quantity compared with a DIS structure function, a diffractive cross
section, or a forward-production bin is a tested color-singlet QCD functional
rather than the Wilson-line state by itself.  Fix a compact small-\(x\)
window, a hard scale range \(Q\ge Q_0\), and a test function
\(\varphi\in\mathcal D_{\rm sx}\) for the measured kinematic bin.  Let
\(\mathfrak T_{Q,Y}[\varphi]\) denote the exact tested functional built from
the relevant current product or inclusive/diffractive cross-section operator
in a finite regulator.  The Wilson-line coordinate associated to the same bin
has the form
\begin{equation}
  \mathcal S^{\rm sx}_{Q,Y}[\varphi]
  :=
  \omega_Y\!\left(\Phi^\varphi_{Q,Y}\right),
  \label{eq:qcd-small-x-wilson-line-observable-coordinate}
\end{equation}
where \(\Phi^\varphi_{Q,Y}\) is the gauge-invariant cylinder-limit Wilson-line
observable produced by the chosen high-energy operator product expansion,
impact factor, rapidity subtraction, and renormalization scheme.  For a
virtual photon this impact factor contains the hard \(q\bar q\) and higher
Fock-state fluctuation weights; for a hadronic final state it contains the
corresponding jet, fragmentation, or diffractive projection.  The transverse
Wilson-line evolution supplies only the target-side rapidity dependence in
this coordinate.

A controlled small-\(x\) prediction for the tested observable therefore
consists of:
\begin{enumerate}
  \item the exact finite-regulator definition of
  \(\mathfrak T_{Q,Y}[\varphi]\) in the measured channel;
  \item an impact-factor matching map from the exact color-singlet operator to
  \(\Phi^\varphi_{Q,Y}\), including virtual corrections, real-emission
  subtractions, and scheme conventions;
  \item a rapidity-factorization statement in which the separation scale
  dependence cancels between the impact factor and JIMWLK evolution in the
  declared topology;
  \item the projective Wilson-line state limit of
  \eqref{eq:qcd-jimwlk-projective-weak-limit-budget} on the cylinder
  representatives entering \(\Phi^\varphi_{Q,Y}\);
  \item an evolution and truncation estimate for the retained finite
  JIMWLK/Balitsky subsystem, with a separate closure residual when BK or a
  moment truncation is used;
  \item a power-remainder estimate for finite \(Q\), finite target size,
  finite rapidity interval, endpoint contamination, and omitted final-state
  measurements.
\end{enumerate}
The finite BFKL eigenvalue, the compact Wilson-line diffusion calculation, and
the BK closure estimate are entries in this list.  They become physics
predictions for the measured functional through the impact-factor map and the
residual controls above.
\end{controlledapproximation}

\begin{controlledapproximation}[Leading small-\(x\) inclusive DIS dipole channel]
\label{ca:qcd-small-x-leading-dis-dipole-channel}
For the inclusive virtual-photon channel the preceding abstract coordinate can
be made explicit at the leading \(q\bar q\) impact-factor level.  Write
\[
  \mathbf x=\mathbf b+\frac{\mathbf r}{2},
  \qquad
  \mathbf y=\mathbf b-\frac{\mathbf r}{2},
  \qquad
  \epsilon_f^2(z,Q):=z(1-z)Q^2+m_f^2 ,
\]
and let \(K_\nu\) denote the modified Bessel function.  In the usual dipole
normalization, following the virtual-photon dipole representation of
\cite{NikolaevZakharov1991,Kovchegov1999SmallXF2}, the squared photon
wave-function weights are
\begin{align}
  W_T^{(f)}(z,r;Q)
  &:=
  \frac{N_c\alpha_{\rm em}e_f^2}{2\pi^2}
  \left(
    [z^2+(1-z)^2]\epsilon_f^2 K_1^2(\epsilon_f r)
    +m_f^2K_0^2(\epsilon_f r)
  \right),
  \label{eq:qcd-small-x-photon-wavefunction-transverse}
\\
  W_L^{(f)}(z,r;Q)
  &:=
  \frac{N_c\alpha_{\rm em}e_f^2}{2\pi^2}
  \left(
    4Q^2z^2(1-z)^2K_0^2(\epsilon_f r)
  \right).
  \label{eq:qcd-small-x-photon-wavefunction-longitudinal}
\end{align}
The leading dipole coordinate for a finite tested bin
\(\varphi\in C_c^\infty((0,1)\times\mathbb R^2_{\mathbf r}
\times\mathbb R^2_{\mathbf b})\) is
\begin{align}
  \mathcal F_i^{\rm dip}[Q,Y_t;\varphi]
  &:=
  2\sum_f
  \int_0^1\dd z
  \int_{\mathbb R^2}\dd^2\mathbf r
  \int_{\mathbb R^2}\dd^2\mathbf b\,
  \varphi(z,\mathbf r,\mathbf b)\,
  W_i^{(f)}(z,r;Q)\,
  \notag
\\
  &\quad{}\times
  N_{Y_t}
  \left(
    \mathbf b+\frac{\mathbf r}{2},
    \mathbf b-\frac{\mathbf r}{2}
  \right),
  \qquad i=T,L .
  \label{eq:qcd-small-x-leading-dis-dipole-functional}
\end{align}
Multiplication by the conventional factor \(Q^2/(4\pi^2\alpha_{\rm em})\)
turns the corresponding inclusive virtual-photon cross section into
\(F_T\) or \(F_L\).  Equation
\eqref{eq:qcd-small-x-leading-dis-dipole-functional} is therefore the first
nontrivial instance of \(\Phi^\varphi_{Q,Y}\): the Wilson-line state is
tested only after convolution with the physical photon fluctuation kernel and
the measured-bin function.

The rapidity split is part of the datum.  Let \(Y_s\) be the target rapidity
assigned to the Wilson-line state and let \(Y-Y_s\) be the projectile
rapidity assigned to the impact factor.  The factorized coordinate has the
form
\[
  \mathcal F_i(Y_s)
  =
  \omega_{Y_s}
  \left(
    \Phi_{i,Q,Y-Y_s}^{\varphi}
  \right).
\]
Using the weak JIMWLK equation gives
\begin{equation}
  \frac{\dd}{\dd Y_s}\mathcal F_i(Y_s)
  =
  \omega_{Y_s}
  \left(
    \mathcal H_{Y_s}\Phi_{i,Q,Y-Y_s}^{\varphi}
    -
    \partial_{Y_{\rm p}}
    \Phi_{i,Q,Y_{\rm p}}^{\varphi}\big|_{Y_{\rm p}=Y-Y_s}
  \right).
  \label{eq:qcd-small-x-dis-rapidity-split-derivative}
\end{equation}
Thus independence of the artificial separation scale is not a slogan: at the
retained logarithmic order the projectile impact-factor subtraction must obey
the same cylinder generator as the target evolution, with the opposite
allocation of rapidity.  If the two sides differ by a cylinder observable
\(R^{\rm rap}_{i,Q,Y_s,\varphi}\), the rapidity residual in the measured DIS
bin is the integral of
\(\omega_{Y_s}(R^{\rm rap}_{i,Q,Y_s,\varphi})\) over the allowed separation
interval.

The BK or finite-moment error is propagated through the same physical kernel.
If a dipole approximation \(\widetilde N_{Y_t}\) satisfies
\[
  |N_{Y_t}(\mathbf b+\mathbf r/2,\mathbf b-\mathbf r/2)
    -\widetilde N_{Y_t}(\mathbf b+\mathbf r/2,\mathbf b-\mathbf r/2)|
  \leq e_N(Y_t;\mathbf r,\mathbf b),
\]
then
\begin{equation}
  \left|
    \mathcal F_i^{\rm dip}[Q,Y_t;\varphi]
    -
    \widetilde{\mathcal F}_i^{\rm dip}[Q,Y_t;\varphi]
  \right|
  \leq
  2\sum_f
  \int\dd z\,\dd^2\mathbf r\,\dd^2\mathbf b\,
  |\varphi|\,W_i^{(f)}(z,r;Q)\,
  e_N(Y_t;\mathbf r,\mathbf b).
  \label{eq:qcd-small-x-dis-bk-error-propagation}
\end{equation}
In particular, a uniform finite-cell BK estimate such as
\eqref{eq:qcd-finite-bk-closure-error-bound} enters the measured structure
function multiplied by the photon-kernel norm of the tested bin, not as an
unweighted abstract residual.

Finally the endpoint and finite-\(Q\) pieces are separate estimates.  For a
regulated retained bin
\[
  \delta\le z\le1-\delta,\qquad
  \rho\le r\le R,\qquad
  \mathbf b\in B ,
\]
the discarded endpoint strips have the explicit spin-factor primitives
\begin{align}
  \int_0^\delta [z^2+(1-z)^2]\,\dd z
  &=
  \delta-\delta^2+\frac23\delta^3
  \le \delta,
  \label{eq:qcd-small-x-transverse-endpoint-spin-bound}
\\
  \int_0^\delta 4z^2(1-z)^2\,\dd z
  &=
  \frac43\delta^3-2\delta^4+\frac45\delta^5
  \le \frac43\delta^3 ,
  \qquad 0\le\delta\le\frac12 .
  \label{eq:qcd-small-x-longitudinal-endpoint-spin-bound}
\end{align}
On the retained bin the Bessel products appearing in \(W_i^{(f)}\) are bounded
by constants depending on \(Q,\delta,\rho,R\).  On the discarded endpoint
strips one must either keep the corresponding tail as an explicit residual or
use the relevant mass, collinear regulator, or integrable massless estimates
for the combinations \(\epsilon_f^2K_1^2(\epsilon_f r)\) and
\(z^2(1-z)^2K_0^2(\epsilon_f r)\).  With those constants, unitarity
\(0\le N_Y\le1\), and the impact-parameter support of \(\varphi\), the
displayed primitives give a finite endpoint remainder for the retained
observable.  The small-\(r\) complement is controlled instead by the
color-transparency input
\(N_Y(\mathbf b+\mathbf r/2,\mathbf b-\mathbf r/2)\le C_B(Y)r^2\) on the
target support; without such an input the dipole formula is only a formal
convolution near \(r=0\).  This is the sense in which the leading DIS dipole
formula is a measured-observable approximation with declared impact-factor,
rapidity, BK/JIMWLK, and finite-\(Q\) remainders, not just a statement about
Wilson-line geometry in transverse coordinate space.
\end{controlledapproximation}

\begin{proposition}[Measured small-\(x\) residual budget]
\label{prop:qcd-small-x-measured-residual-budget}
In the setting of
Controlled Approximation~\ref{ca:qcd-small-x-measured-observable-map}, suppose
the chosen finite approximation
\(\mathcal A^{(m,n)}_{Q,Y}[\varphi]\) to
\(\mathcal S^{\rm sx}_{Q,Y}[\varphi]\) satisfies
\begin{equation}
\begin{aligned}
  \mathfrak T_{Q,Y}[\varphi]
  =
  \mathcal A^{(m,n)}_{Q,Y}[\varphi]
  &+\Delta_{\rm IF}[\varphi]
  +\Delta_{\rm rap}[\varphi]
  +\Delta_{\rm proj}[\varphi]
\\
  &+\Delta_{\rm evol}[\varphi]
  +\Delta_{\rm clos}[\varphi]
  +\Delta_{\rm pwr}[\varphi],
\end{aligned}
\label{eq:qcd-small-x-measured-residual-decomposition}
\end{equation}
where the terms denote, respectively, impact-factor matching, rapidity
subtraction, projective regulator removal, finite evolution/truncation,
closure, and power remainders.  If, on the chosen window,
\begin{equation}
  |\Delta_a[\varphi]|
  \leq
  \epsilon_a(Q_0,Y_*)\,\|\varphi\|_{\mathcal D_{\rm sx}},
  \qquad
  a\in
  \{\mathrm{IF},\mathrm{rap},\mathrm{proj},
    \mathrm{evol},\mathrm{clos},\mathrm{pwr}\},
  \label{eq:qcd-small-x-measured-residual-component-bounds}
\end{equation}
then
\begin{equation}
  \left|
    \mathfrak T_{Q,Y}[\varphi]
    -
    \mathcal A^{(m,n)}_{Q,Y}[\varphi]
  \right|
  \leq
  \left(
    \epsilon_{\rm IF}
    +\epsilon_{\rm rap}
    +\epsilon_{\rm proj}
    +\epsilon_{\rm evol}
    +\epsilon_{\rm clos}
    +\epsilon_{\rm pwr}
  \right)
  \|\varphi\|_{\mathcal D_{\rm sx}} .
  \label{eq:qcd-small-x-measured-residual-budget-bound}
\end{equation}
For full JIMWLK evolution one may set
\(\epsilon_{\rm clos}=0\) and keep the evolution error in
\(\epsilon_{\rm evol}\); for a BK or finite-moment projection,
\(\epsilon_{\rm clos}\) is the measured-channel image of the connected
Wilson-line residual such as \(E^{\rm BK}_{\mathbf x\mathbf y}\).
\end{proposition}

\begin{proof}
Subtract the finite approximant in
\eqref{eq:qcd-small-x-measured-residual-decomposition} and apply the triangle
inequality with
\eqref{eq:qcd-small-x-measured-residual-component-bounds}.  The projective
term is controlled by the cylinder weak-equation ledger
\eqref{eq:qcd-jimwlk-projective-weak-limit-budget} after inserting the
cylinder representative of \(\Phi^\varphi_{Q,Y}\).  The impact-factor and
power terms are separate QCD statements about the measured color-singlet
operator, while the finite BFKL/Balitsky/JIMWLK calculations constrain their
corresponding coordinates inside the displayed sum.
\end{proof}

\paragraph{Conformal covariance of the transverse kernel.}
The measure
\[
  \dd^2\mathbf z\,M_{\mathbf x\mathbf y\mathbf z}
\]
is invariant under translations, rotations, uniform scalings, and inversion
\(\mathbf u\mapsto\mathbf u/|\mathbf u|^2\), whenever the transformed points
avoid the inversion center.  Thus the leading fixed-coupling transverse
kernel is covariant under the real Mobius transformations generated by these
maps.

Translations and rotations preserve all distances and \(\dd^2\mathbf z\).
Under the scaling \(\mathbf u\mapsto\lambda\mathbf u\), the kernel
\(M_{\mathbf x\mathbf y\mathbf z}\) scales as \(\lambda^{-2}\), while
\(\dd^2\mathbf z\) scales as \(\lambda^2\).

For inversion, write \(\mathbf u'=\mathbf u/|\mathbf u|^2\).  The Euclidean
distance and area element transform as
\[
  |\mathbf u'-\mathbf v'|^2
  =
  \frac{|\mathbf u-\mathbf v|^2}
       {|\mathbf u|^2|\mathbf v|^2},
  \qquad
  \dd^2\mathbf z'
  =
  \frac{\dd^2\mathbf z}{|\mathbf z|^4}.
\]
Substituting these identities gives
\[
  M_{\mathbf x'\mathbf y'\mathbf z'}\,\dd^2\mathbf z'
  =
  \frac{
    r_{xy}^2/(|\mathbf x|^2|\mathbf y|^2)
  }{
    r_{xz}^2/(|\mathbf x|^2|\mathbf z|^2)\,
    r_{zy}^2/(|\mathbf z|^2|\mathbf y|^2)
  }
  \frac{\dd^2\mathbf z}{|\mathbf z|^4}
  =
  M_{\mathbf x\mathbf y\mathbf z}\,\dd^2\mathbf z.
\]
This proves the stated covariance.

\begin{proposition}[Mellin eigenvalue of the leading BFKL kernel]
\label{prop:qcd-bfkl-mellin-eigenvalue}
Let \(0<\operatorname{Re}\gamma<1\).  For the power
\[
  f_\gamma(\mathbf x,\mathbf y)
  =
  \bigl(\mu_\perp^2 r_{xy}^2\bigr)^\gamma ,
\]
define the subtracted transverse integral by analytic regularization as in
the proof below.  Then
\begin{equation}
  \frac1{2\pi}\int_{\mathbb R^2}\dd^2\mathbf z\,
  M_{\mathbf x\mathbf y\mathbf z}
  \bigl[
    f_\gamma(\mathbf x,\mathbf z)
    +f_\gamma(\mathbf z,\mathbf y)
    -f_\gamma(\mathbf x,\mathbf y)
  \bigr]
  =
  \chi(\gamma)f_\gamma(\mathbf x,\mathbf y),
  \label{eq:qcd-bfkl-eigenvalue-equation}
\end{equation}
where
\begin{equation}
  \chi(\gamma)
  =
  2\psi(1)-\psi(\gamma)-\psi(1-\gamma),
  \qquad
  \psi(u):=\frac{\dd}{\dd u}\log\Gamma(u).
  \label{eq:qcd-bfkl-characteristic-function}
\end{equation}
\end{proposition}

\begin{proof}
By translation and scaling it suffices to set
\(\mathbf y=0\), \(r_{xy}=1\), and \(\mu_\perp=1\).  The singular terms in
the three brackets are regulated in \(d=2-2\epsilon\) transverse dimensions.
The Euclidean beta integral
\begin{equation}
  \int\dd^d\mathbf z\,
  \frac1{(\mathbf z^2)^a((\mathbf z-\mathbf x)^2)^b}
  =
  \pi^{d/2}
  \frac{
    \Gamma(d/2-a)\Gamma(d/2-b)\Gamma(a+b-d/2)
  }{
    \Gamma(a)\Gamma(b)\Gamma(d-a-b)
  }
  (\mathbf x^2)^{d/2-a-b}
  \label{eq:qcd-bfkl-beta-integral}
\end{equation}
is first valid in its convergence domain and then analytically continued.
Applying \eqref{eq:qcd-bfkl-beta-integral} to the three pieces gives
\[
\begin{aligned}
  J_\epsilon(\gamma)
  &=
  \pi^{1-\epsilon}
  \frac{
    \Gamma(\gamma-\epsilon)\Gamma(-\epsilon)
    \Gamma(1-\gamma+\epsilon)
  }{
    \Gamma(1-\gamma)\Gamma(\gamma-2\epsilon)
  },\\
  J_\epsilon(1-\gamma)
  &=
  \pi^{1-\epsilon}
  \frac{
    \Gamma(1-\gamma-\epsilon)\Gamma(-\epsilon)
    \Gamma(\gamma+\epsilon)
  }{
    \Gamma(\gamma)\Gamma(1-\gamma-2\epsilon)
  },\\
  J_\epsilon(0)
  &=
  \pi^{1-\epsilon}
  \frac{\Gamma(-\epsilon)^2\Gamma(1+\epsilon)}
       {\Gamma(-2\epsilon)} .
\end{aligned}
\]
The regulated subtracted integral is
\[
  J_\epsilon(\gamma)+J_\epsilon(1-\gamma)-J_\epsilon(0).
\]
Using
\[
  \Gamma(u+\delta)
  =
  \Gamma(u)\bigl(1+\delta\psi(u)+O(\delta^2)\bigr)
\]
and the corresponding expansion at the pole \(u=0\), the \(1/\epsilon\)
terms cancel and the finite part is
\[
  2\pi\,[2\psi(1)-\psi(\gamma)-\psi(1-\gamma)].
\]
Dividing by \(2\pi\) proves
\eqref{eq:qcd-bfkl-eigenvalue-equation}.
\end{proof}

The leading fixed-coupling kernel in
\eqref{eq:qcd-coordinate-bfkl-equation} has
\[
  \chi\!\left(\frac12\right)=4\log2,
  \qquad
  \chi\!\left(\frac12+\ii\nu\right)
  =
  4\log2-14\zeta(3)\nu^2+O(\nu^4).
\]
These values follow from
\[
  \psi\!\left(\frac12\right)=-\gamma_{\rm E}-2\log2,
  \qquad
  \psi''\!\left(\frac12\right)=-14\zeta(3),
\]
inserted into \eqref{eq:qcd-bfkl-characteristic-function}; in particular
\(\chi'(\frac12)=0\) and \(\chi''(\frac12)=28\zeta(3)\).  A Mellin component
near \(\gamma=\frac12\) therefore evolves as
\[
  \exp\!\left[
    \bar\alpha_{\rm s}\,4\log2\,Y
    -\bar\alpha_{\rm s}\,14\zeta(3)\nu^2Y
    +O(\bar\alpha_{\rm s}\nu^4Y)
  \right]
\]
inside the fixed-coupling leading-logarithmic setup.

\begin{controlledapproximation}[Status of small-\(x\) resummation]
\label{ca:qcd-small-x-status}
A small-\(x\) resummation used for a QCD observable must state:
the regulated Wilson-line operator being evolved; the rapidity variable and
subtraction scheme; the impact factor connecting the Wilson-line matrix
element to the measured process; whether running coupling, energy-scale,
collinear-improvement, saturation, or finite-\(N_c\) hierarchy effects are
included; and the topology of the remainder estimate.  The BFKL kernel in
Proposition~\ref{prop:qcd-bfkl-mellin-eigenvalue} is a theorem about the
leading transverse integral.  Its physical use in high-energy hadron
scattering additionally requires the Wilson-line operator definition, impact
factor, rapidity-factorization scheme, and unitarization or saturation
framework appropriate to the observable.
\end{controlledapproximation}
