\chapter{The BV Master Formalism for Gauge Effective Actions}
\label{ch:bv-master-formalism}
\ConventionsMixed


\section{Why Antifields Are Introduced}

The BRST chapter introduced sources \(K^{a\mu}\) and \(L^a\) for the nonlinear
variations \(sA_\mu^a\) and \(sc^a\).  Their role is to serve as coordinates
dual to the BRST variations, thereby making the Slavnov--Taylor identity a
functional equation for the 1PI effective action.

The Batalin--Vilkovisky formalism is the intrinsic version of this
construction.  It assigns a dual coordinate, called an antifield, to every
field in the gauge-fixed or ungauge-fixed complex.  The antibracket pairs a
field with its antifield, and the master equation is the single equation that
encodes gauge invariance, BRST nilpotency, gauge fixing, and the compatible
renormalization identities.

The framework in this chapter is perturbative and local.  We work with a
regulated field theory or, equivalently for algebraic purposes, with the
graded algebra of local functionals in finitely many jets of fields.  Infinite
dimensional functional derivatives are therefore shorthand for the
corresponding regulated or local-jet operation.  When an identity involves the
BV Laplacian, the regulator and measure are part of the data.

This chapter supplies the continuum gauge-theory framework for effective
actions.  A Wilsonian or 1PI effective action for a gauge theory is a
gauge-compatible object only after the gauge symmetry has been encoded in a
BRST/BV complex, or after a lattice regulator has supplied a finite
gauge-invariant integral over link variables.  In the BV route the finite
cutoff data are the odd symplectic field-antifield space, a compatible
half-density, the regulated BV Laplacian, the master action, and the
coarse-graining cycle.  In the lattice route the finite cutoff data are the
compact link-variable space, Haar measure, gauge action, gauge-invariant
observables, and blocking or scaling map.  Later gauge-theory Wilsonian
claims use one of these two data types and state the comparison map between
them when both appear.

\section{Fields, Antifields, and Gradings}

Let \(\{\Phi^A\}\) denote the full set of fields in the gauge-theory complex.
The index \(A\) includes discrete labels, spacetime indices, Lie-algebra
indices, and the spacetime point in condensed notation.  Examples are
\[
  A_\mu^a,\qquad c^a,\qquad \bar c^a,\qquad B^a,
  \qquad \psi,\qquad \bar\psi .
\]
Each field has a Grassmann parity
\[
  \epsilon_A\in\mathbb Z/2\mathbb Z
\]
and a ghost number \(\operatorname{gh}(\Phi^A)\).  The antifield paired with
\(\Phi^A\) is denoted by \(\Phi^*_A\) and is assigned
\[
  \epsilon(\Phi^*_A)=\epsilon_A+1\pmod 2,
  \qquad
  \operatorname{gh}(\Phi^*_A)=-1-\operatorname{gh}(\Phi^A).
\]
Thus antifields have opposite parity and complementary ghost number.  For the
Yang--Mills nonminimal sector,
\[
\begin{array}{c|cccc}
  \Phi^A & A_\mu & c & \bar c & B\\ \hline
  \operatorname{gh}(\Phi^A) & 0 & 1 & -1 & 0\\
  \operatorname{gh}(\Phi^*_A) & -1 & -2 & 0 & -1
\end{array}
\]
and the parities are reversed in the antifield row.

Let \(\mathcal F_{\rm BV}\) be the graded-commutative algebra of local
functionals in \(\Phi^A,\Phi^*_A\) and finitely many of their spacetime
derivatives, modulo integration by parts when integrated local functionals are
being compared.  A functional \(F\in\mathcal F_{\rm BV}\) has parity
\(|F|\) and ghost number \(\operatorname{gh}(F)\) when it is homogeneous.
Left and right variational derivatives are defined by
\[
  \delta F
  =
  \int
  \delta\Phi^A\,
  \frac{\delta^L F}{\delta\Phi^A}
  =
  \int
  \frac{\delta^R F}{\delta\Phi^A}\,
  \delta\Phi^A ,
\]
and similarly for antifields.  The distinction is necessary because odd
variables do not commute with the variation symbols.

For a local functional
\[
  F[\Phi]=\int_M f(x,\Phi,\partial\Phi,\ldots,\partial^r\Phi)\,\dd^D x ,
\]
the left Euler--Lagrange derivative is the coefficient of
\(\delta\Phi^A\) after moving derivatives off the variation:
\begin{equation}
  \frac{\delta^L F}{\delta\Phi^A}
  =
  \sum_{|\alpha|\leq r}
  (-\partial)_\alpha
  \frac{\partial^L f}
       {\partial(\partial_\alpha\Phi^A)} .
  \label{eq:bv-euler-operator}
\end{equation}
Here \(\alpha\) is a multi-index and
\((-\partial)_\alpha=(-1)^{|\alpha|}\partial_\alpha\).  Boundary conditions or
compact support are part of the definition whenever the integration by parts
is used.  In a finite regulator \(\partial_\alpha\Phi^A\) is replaced by the
corresponding finite set of regulated coordinates, and
\eqref{eq:bv-euler-operator} becomes the ordinary graded partial derivative.
The ghost number and parity of the variational derivative are
\[
  \operatorname{gh}\!\left(\frac{\delta^L F}{\delta\Phi^A}\right)
  =
  \operatorname{gh}(F)-\operatorname{gh}(\Phi^A),
  \qquad
  \left|\frac{\delta^L F}{\delta\Phi^A}\right|
  =
  |F|+\epsilon_A .
\]
These bookkeeping rules are what make the BV bracket below homogeneous of
parity one and ghost number one.

\section{The Antibracket}

\begin{definition}[BV antibracket]
For homogeneous functionals \(F,G\in\mathcal F_{\rm BV}\), the BV
antibracket is
\begin{equation}
  (F,G)
  =
  \int
  \left[
  \frac{\delta^R F}{\delta\Phi^A}
  \frac{\delta^L G}{\delta\Phi^*_A}
  -
  \frac{\delta^R F}{\delta\Phi^*_A}
  \frac{\delta^L G}{\delta\Phi^A}
  \right]\dd^D x ,
  \label{eq:bv-antibracket}
\end{equation}
with summation over all field labels \(A\).  The spacetime integral is absent
when the same formula is applied to a finite-dimensional regulator.
\end{definition}

The bracket has parity and ghost number
\[
  |(F,G)|=|F|+|G|+1,
  \qquad
  \operatorname{gh}(F,G)=\operatorname{gh}(F)+\operatorname{gh}(G)+1.
\]
It satisfies
\[
  (F,G)
  =
  -(-1)^{(|F|+1)(|G|+1)}(G,F)
\]
and the odd Jacobi identity
\[
  (F,(G,H))
  =
  ((F,G),H)
  +
  (-1)^{(|F|+1)(|G|+1)}(G,(F,H)).
\]
It is an odd biderivation.  Explicitly,
\[
  (F,GH)
  =
  (F,G)H
  +
  (-1)^{(|F|+1)|G|}G(F,H),
\]
and
\[
  (FG,H)
  =
  F(G,H)
  +
  (-1)^{|G|(|H|+1)}(F,H)G .
\]
The elementary brackets are
\[
  (\Phi^A(x),\Phi^*_B(y))=\delta^A{}_B\,\delta^{(D)}(x-y),
  \qquad
  (\Phi^A,\Phi^B)=0,
  \qquad
  (\Phi^*_A,\Phi^*_B)=0,
\]
with the delta distribution understood in the same regulated or distributional
sense as the fields themselves.

\begin{proposition}[Odd symplectic origin of the antibracket]
\label{prop:bv-antibracket-odd-symplectic-origin}
At finite regulator, let
\[
  \omega=\delta\Phi^*_A\,\delta\Phi^A
\]
be the odd symplectic form on the graded vector space with coordinates
\((\Phi^A,\Phi^*_A)\).  For a homogeneous function \(F\), define its
Hamiltonian vector field \(X_F\) by
\[
  \iota_{X_F}\omega=\delta F .
\]
Then the bracket \((F,G):=X_FG\) is exactly
\eqref{eq:bv-antibracket}.  It satisfies the elementary brackets,
graded antisymmetry, the odd Jacobi identity, and the two biderivation
identities displayed above.
\end{proposition}

\begin{proof}
In Darboux coordinates the equation
\(\iota_{X_F}\omega=\delta F\) gives
\[
  X_F
  =
  \frac{\delta^R F}{\delta\Phi^A}
  \frac{\delta^L}{\delta\Phi^*_A}
  -
  \frac{\delta^R F}{\delta\Phi^*_A}
  \frac{\delta^L}{\delta\Phi^A}.
\]
Applying this vector field to \(G\) gives
\eqref{eq:bv-antibracket}.  The elementary brackets follow by applying the
formula to the coordinate functions.  The biderivation identities follow
because \(X_F\) is a graded derivation of parity \(|F|+1\).

For the Jacobi identity, use the defining relation and the closedness of the
constant Darboux form:
\[
  \iota_{[X_F,X_G]}\omega
  =
  \delta(X_FG)
  =
  \delta(F,G).
\]
Thus \([X_F,X_G]=X_{(F,G)}\), with the graded commutator of vector fields.
The graded Jacobi identity for vector-field commutators gives the displayed
odd Jacobi identity for \((\,\cdot\,,\,\cdot\,)\).  For continuum local
functionals, first smear the fields or impose a finite regulator so that the
Darboux coordinate calculation is finite-dimensional.  The variational
derivatives in \eqref{eq:bv-antibracket} are then the limits of the regulated
coordinate derivatives, and the bracket identities pass to the declared
local-functional limit on the common domain where those variational
derivatives exist.
\end{proof}

\section{The Classical Master Equation}

\begin{definition}[Classical BV action]
A classical BV action is an even local functional \(S_{\rm BV}\in
\mathcal F_{\rm BV}\) of ghost number zero satisfying the classical master
equation
\begin{equation}
  \frac12(S_{\rm BV},S_{\rm BV})=0 .
  \label{eq:classical-master-equation}
\end{equation}
It is required to reduce to the classical gauge-invariant action when all
ghosts and antifields are set to zero.
\end{definition}

The associated BV differential is
\[
  s_{\rm BV}F=(F,S_{\rm BV}).
\]

If \(S_{\rm BV}\) satisfies \eqref{eq:classical-master-equation}, then
\[
  s_{\rm BV}^2=0 .
\]
Using the odd Jacobi identity and the evenness of \(S_{\rm BV}\),
\[
  ((H,S_{\rm BV}),S_{\rm BV})
  =
  \frac12(H,(S_{\rm BV},S_{\rm BV}))
\]
for every functional \(H\).  The right hand side vanishes by
\eqref{eq:classical-master-equation}.  Hence
\[
  s_{\rm BV}^2H=((H,S_{\rm BV}),S_{\rm BV})=0 .
\]
Since this holds for every functional \(H\) in the BV algebra, the Hamiltonian
vector field generated by \(S_{\rm BV}\) is a differential.  Thus the
classical master equation is precisely the condition that the BRST/BV
variation squares to zero off shell, before choosing a gauge-fixing
Lagrangian submanifold.

This algebraic implication is the reason the master equation is stronger than
a compact rewriting of a Ward identity.  It constructs a differential on the
space of fields, ghosts, antifields, and local functionals.

\section{Noether Identities and the Minimal BV Complex}

The master equation is constructed from the gauge identities of the classical
action.  Let \(S_0[\phi]\) be a classical action for fields \(\phi^i\), and set
\[
  E_i(\phi):=\frac{\delta S_0}{\delta\phi^i}.
\]
An infinitesimal gauge transformation is a collection of local differential
operators \(R^i{}_\alpha(\phi,\partial)\) acting on gauge parameters
\(\epsilon^\alpha\):
\[
  \delta_\epsilon\phi^i=R^i{}_\alpha\epsilon^\alpha .
\]
Gauge invariance of \(S_0\) is the Noether identity
\begin{equation}
  E_iR^i{}_\alpha=0
  \label{eq:bv-noether-identity}
\end{equation}
after moving derivatives off \(\epsilon^\alpha\).  Condensed notation includes
the spacetime integration and the signs from integration by parts.

The first BV variables are the ghosts \(c^\alpha\), the field antifields
\(\phi^*_i\), and the ghost antifields \(c^*_\alpha\).  Their ghost numbers
are
\[
  \operatorname{gh}(\phi^i)=0,\quad
  \operatorname{gh}(c^\alpha)=1,\quad
  \operatorname{gh}(\phi^*_i)=-1,\quad
  \operatorname{gh}(c^*_\alpha)=-2.
\]
The Koszul--Tate part of the BV differential resolves the equations of
motion:
\[
  \delta_{\rm KT}\phi^i=0,\qquad
  \delta_{\rm KT}c^\alpha=0,\qquad
  \delta_{\rm KT}\phi^*_i=E_i,\qquad
  \delta_{\rm KT}c^*_\alpha=R^i{}_\alpha\phi^*_i .
\]
Its square on \(c^*_\alpha\) is
\[
  \delta_{\rm KT}^2 c^*_\alpha
  =
  R^i{}_\alpha E_i,
\]
which vanishes by \eqref{eq:bv-noether-identity}.  Thus the Noether identity
is the first exactness condition in the antifield direction.

For an irreducible off-shell closed gauge algebra,
\[
  [R_\alpha,R_\beta]^i
  =
  R^i{}_\gamma C^\gamma{}_{\alpha\beta},
\]
with the usual graded signs in the general case, the beginning of the minimal
BV action is
\begin{equation}
  S_{\rm min}
  =
  S_0
  +
  \phi^*_iR^i{}_\alpha c^\alpha
  -
  \frac12 c^*_\gamma C^\gamma{}_{\alpha\beta}c^\alpha c^\beta
  +\cdots .
  \label{eq:bv-general-minimal-action}
\end{equation}
For field-independent structure constants and off-shell closure this displayed
expression is complete.  Field-dependent, open, or reducible gauge systems
require higher antifield-number terms.  Expanding
\(\frac12(S_{\rm min},S_{\rm min})\) by antifield number gives, in order,
the Noether identity \eqref{eq:bv-noether-identity}, the closure relation of
the gauge generators, and the Jacobi relations or their higher homotopy
replacements.  Solving the master equation is therefore the construction of a
single nilpotent differential whose antifield expansion records the complete
gauge identity complex.

\section{Minimal Yang--Mills BV Action}

For a gauge theory whose BRST transformations close off shell, the minimal BV
action is obtained by coupling antifields to the BRST variations.  With the
Hermitian Yang--Mills conventions of
Chapter~\ref{ch:vol2-classical-yang-mills-matter},
\([t^a,t^b]=\ii f^c{}_{ab}t^c\), so the output Lie-algebra index of
\(f^c{}_{ab}\) is written first.  In this convention
\[
  sA_\mu^a=(D_\mu c)^a,\qquad
  sc^a=-\frac12 f^a{}_{bc}c^b c^c .
\]
The minimal BV action for the gauge and ghost sector is
\begin{equation}
  S_{\rm min}
  =
  S_{\rm YM}[A]
  +
  \int_M
  \left[
  A^{*\mu}_a(D_\mu c)^a
  -
  \frac12 c^*_a f^a{}_{bc}c^b c^c
  \right]\dd^D x .
  \label{eq:minimal-yang-mills-bv-action}
\end{equation}
Matter fields are included by adding their classical action and antifield
couplings
\[
  \int_M
  \left[
  \psi^*\,(\ii c^a t_R^a\psi)
  +
  (-\ii\bar\psi\,c^a t_R^a)\,\bar\psi^*
  \right]\dd^D x ,
\]
with the order of odd factors fixed by the left/right derivative convention.

The master equation for \eqref{eq:minimal-yang-mills-bv-action} is the
combined statement that \(S_{\rm YM}\) is gauge invariant and that the BRST
transformations are nilpotent.  Let
\[
  R_\mu^a(A,c):=(D_\mu c)^a,
  \qquad
  R_c^a(c):=-\frac12 f^a{}_{bc}c^b c^c,
\]
and write
\[
  S_{\rm min}=S_{\rm YM}+S_1,
  \qquad
  S_1=\int_M\left(A^{*\mu}_a R_\mu^a+c^*_aR_c^a\right)\dd^D x .
\]
Give each antifield antifield number \(1\) and each field antifield number
\(0\).  Since \(S_{\rm YM}\) is independent of antifields and \(S_1\) is
linear in antifields,
\[
  \frac12(S_{\rm min},S_{\rm min})
  =
  (S_{\rm YM},S_1)+\frac12(S_1,S_1),
\]
and the two displayed terms have antifield numbers \(0\) and \(1\),
respectively.  The antifield-number-zero component is
\[
  (S_{\rm YM},S_1)
  =
  \int_M
  \frac{\delta^R S_{\rm YM}}{\delta A_\mu^a}
  R_\mu^a\,\dd^D x
  =
  \int_M
  \frac{\delta S_{\rm YM}}{\delta A_\mu^a}
  (D_\mu c)^a\,\dd^D x .
\]
Gauge invariance of \(S_{\rm YM}\) means that this expression vanishes when
the odd ghost \(c\) is replaced by an arbitrary even infinitesimal gauge
parameter \(\zeta\).  Because the identity is algebraic in the parameter and
its derivatives, the same coefficient identity gives the vanishing for \(c\).

For the antifield-number-one component, the antibracket calculation gives
\[
  \frac12(S_1,S_1)
  =
  \int_M
  \left[
    A^{*\mu}_a\,sR_\mu^a
    +
    c^*_a\,sR_c^a
  \right]\dd^D x ,
\]
where \(s\) acts on functions of \(A,c\) by
\[
  sX
  =
  \int_M
  \left[
    R_\mu^a\frac{\delta_L X}{\delta A_\mu^a}
    +
    R_c^a\frac{\delta_L X}{\delta c^a}
  \right]\dd^D x .
\]
The two coefficients are explicitly
\[
\begin{aligned}
  sR_\mu^a
  &=
  D_\mu R_c^a
  +
  f^a{}_{bc}R_\mu^b c^c
\\
  &=
  -\frac12 f^a{}_{bc}D_\mu(c^b c^c)
  +
  f^a{}_{bc}(D_\mu c)^b c^c
  =
  0,
\end{aligned}
\]
where \(D_\mu\) is an even derivation and the two terms in
\(D_\mu(c^b c^c)\) are equal after antisymmetry of \(f^a{}_{bc}\) and the odd
commutation of \(c\) with \(D_\mu c\).  Similarly,
\[
\begin{aligned}
  sR_c^a
  &=
  -\frac12 f^a{}_{bc}
  \left(R_c^b c^c-c^b R_c^c\right)
\\
  &=
  -\frac1{12}
  \left(
    f^a{}_{be}f^e{}_{cd}
    +
    f^a{}_{ce}f^e{}_{db}
    +
    f^a{}_{de}f^e{}_{bc}
  \right)c^b c^c c^d
  =
  0
\end{aligned}
\]
by the Jacobi identity.  There are no higher antifield-number components for
this off-shell closed Yang--Mills algebra.  Thus
\(\frac12(S_{\rm min},S_{\rm min})=0\) is exactly the pair of statements
\(sS_{\rm YM}=0\) and \(s^2A=s^2c=0\).

\begin{remark}[Off-shell closed and open gauge algebras]
The preceding construction used a special property of Yang--Mills theory.  Let
\(\Phi^i\) denote the classical fields, let
\[
  E_i(\Phi):=\frac{\delta S_0}{\delta\Phi^i}
\]
be the Euler--Lagrange derivative of the classical gauge-invariant action, and
write an infinitesimal gauge variation as
\[
  \delta_\epsilon\Phi^i=R^i{}_\alpha(\Phi)\epsilon^\alpha .
\]
Here \(i\) and \(\alpha\) are condensed indices, including spacetime position
and possible derivatives acting on the gauge parameter.  The algebra is
off-shell closed when the commutator of gauge vector fields satisfies an
identity on field space,
\begin{equation}
  R^j{}_\beta\frac{\delta R^i{}_\alpha}{\delta\Phi^j}
  -
  R^j{}_\alpha\frac{\delta R^i{}_\beta}{\delta\Phi^j}
  =
  R^i{}_\gamma C^\gamma{}_{\alpha\beta}(\Phi),
  \label{eq:bv-off-shell-closed-algebra}
\end{equation}
with no use of \(E_i=0\).  Yang--Mills theory has
\(C^\gamma{}_{\alpha\beta}=f^\gamma{}_{\alpha\beta}\), in the local
Lie-algebra notation of the BRST chapter.  Then the minimal action is linear
in antifields, apart from the ghost antifield term encoding
\eqref{eq:bv-off-shell-closed-algebra}, and the master equation is precisely
the gauge-invariance and Jacobi-identity calculation displayed above.

An open algebra closes only modulo the equations of motion.  For bosonic
gauge parameters and bosonic fields, the corresponding commutator identity is
\begin{equation}
  R^j{}_\beta\frac{\delta R^i{}_\alpha}{\delta\Phi^j}
  -
  R^j{}_\alpha\frac{\delta R^i{}_\beta}{\delta\Phi^j}
  =
  R^i{}_\gamma C^\gamma{}_{\alpha\beta}(\Phi)
  +
  M^{ij}{}_{\alpha\beta}(\Phi)E_j(\Phi),
  \label{eq:bv-open-algebra}
\end{equation}
with the standard Koszul signs inserted in theories with fermionic fields or
fermionic gauge parameters.  The coefficient
\(M^{ij}{}_{\alpha\beta}\) is graded antisymmetric in the pair
\((i,j)\) in the sense required by the Noether identities.  If one couples
antifields only linearly to the naive BRST variations, the antibracket
\(\frac12(S,S)\) contains a term proportional to
\(\Phi^*_i M^{ij}{}_{\alpha\beta}E_j c^\alpha c^\beta\).  This is zero after
imposing the classical equations of motion, but it is not zero as an identity
on the field algebra.  The BV solution cancels it by adding higher
antifield-number terms.  In the same ungraded sign convention, the first such
term is a quadratic antifield term of the form
\[
  \frac14\int \Phi^*_i\Phi^*_j
  M^{ij}{}_{\alpha\beta}(\Phi)c^\alpha c^\beta ,
\]
with signs fixed by the parity convention.  Further terms are then fixed
recursively by the single equation
\(\frac12(S_{\rm BV},S_{\rm BV})=0\).  Thus BV constructs, in open algebras,
an off-shell
nilpotent differential on the extended field-antifield algebra.

Reducible symmetries require additional ghosts.  If
\[
  R^i{}_\alpha Z^\alpha{}_{a_1}=0
\]
identically, or modulo the equations of motion, then the gauge parameters
themselves have gauge redundancies, and one introduces ghosts-for-ghosts
together with their antifields.  Standard examples where the BV framework is
structural rather than optional include supergravity theories without a
complete off-shell auxiliary-field formulation, type-II superstring field
theory, and the Poisson sigma model.  The elementary Faddeev--Popov/BRST
construction in the preceding chapter concerns the off-shell closed
Yang--Mills case.
\end{remark}

\section{Nonminimal Sector and Gauge Fixing}

Gauge fixing is implemented by adding a contractible nonminimal pair and then
choosing a Lagrangian submanifold in the BV field-antifield space.  The
cohomological content is carried by the same BV differential.

For Yang--Mills theory the nonminimal pair is \((\bar c^a,B^a)\), with
\[
  s\bar c^a=B^a,\qquad sB^a=0.
\]
The nonminimal BV term is
\[
  S_{\rm nm}
  =
  \int_M \bar c^*_a B^a\,\dd^D x .
\]
The total classical BV action is
\[
  S_{\rm BV}=S_{\rm min}+S_{\rm nm}.
\]
In a gauge-fixed BV functional integral the variables \(\bar c^a\) and \(B^a\)
belong to the nonminimal integration sector.  The field \(B^a\) is an even
auxiliary integration variable, not an external source.  Its purpose is to
make the nonminimal BRST pair off-shell contractible and to write the
gauge-fixing term before completing the square.

\begin{definition}[Gauge-fixing fermion]
A gauge-fixing fermion is an odd functional \(\Psi[\Phi]\) of ghost number
\(-1\) depending only on the fields.  It determines the Lagrangian
submanifold
\begin{equation}
  \Phi^*_A=\frac{\delta^L\Psi}{\delta\Phi^A}.
  \label{eq:gauge-fixing-lagrangian-submanifold}
\end{equation}
The gauge-fixed action is the restriction
\[
  S_\Psi[\Phi]
  =
  S_{\rm BV}\left[\Phi,\Phi^*=\frac{\delta^L\Psi}{\delta\Phi}\right].
\]
\end{definition}

For covariant Yang--Mills gauges,
\[
  \Psi
  =
  \int_M
  \bar c^a\left(
  \partial^\mu A_\mu^a+\frac{g_{\rm YM}^2\xi}{2}B^a
  \right)\dd^D x .
\]
Substituting \eqref{eq:gauge-fixing-lagrangian-submanifold} into
\(S_{\rm BV}\), and integrating by parts with the same boundary convention as
in the BRST chapter, gives
\[
  S_\Psi
  =
  S_{\rm YM}
  +
  \int_M
  \left[
  B^a\partial^\mu A_\mu^a
  +
  \frac{g_{\rm YM}^2\xi}{2}B^aB^a
  -
  \bar c^a\partial^\mu D_\mu^{ab}(A)c^b
  \right]\dd^D x .
\]
Thus the Faddeev--Popov action is the restriction of the BV action to a
gauge-fixing Lagrangian submanifold.  At fixed regulator, the subsequent
integration over \(B\) is Gaussian:
\[
  \int[DB]\,
  \exp\left\{
  \ii\int_M
  \left[
  B^a\partial^\mu A_\mu^a
  +
  \frac{g_{\rm YM}^2\xi}{2}B^aB^a
  \right]\dd^D x
  \right\}
  =
  \mathcal N_B(\xi)\,
  \exp\left\{
  -\frac{\ii}{2g_{\rm YM}^2\xi}
  \int_M(\partial^\mu A_\mu^a)^2\,\dd^D x
  \right\}.
\]
After this auxiliary integral is performed, \(B\) is no longer a remaining
field variable in the gauge-fixed path integral; retaining it instead is an
equivalent first-order presentation of the same regularized Gaussian factor.
Different gauge choices correspond to different Lagrangian submanifolds.  They
define equivalent physical cohomology when the BV integral has no boundary
contribution and the quantum master equation is preserved.

Figure~\ref{fig:bv-field-antifield-master-map} displays the geometric
content of these formulae.  The BV space is an odd symplectic supermanifold
with Darboux coordinates \((\Phi^A,\Phi^*_A)\).  A gauge-fixing fermion does
not add a new dynamical variable; it chooses a Lagrangian submanifold on which
the antifields are functions of the fields.

\begin{figure}[H]
  \centering
  \resizebox{0.92\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,2.95)
      {BV field--antifield geometry and gauge fixing};

    \begin{scope}[shift={(-2.7,0)}]
      \fill[blue!4] (-4.45,-2.25) rectangle (3.15,2.45);
      \draw[thick] (-4.45,-2.25) rectangle (3.15,2.45);
      \node[anchor=north west,align=left] at (-4.25,2.28)
        {\(\mathcal E_{\rm BV}\)};
      \draw[-{Latex[length=2mm]},thick] (-4.05,-1.75)--(2.75,-1.75)
        node[below] {fields and ghosts \(\Phi^A\)};
      \draw[-{Latex[length=2mm]},thick] (-3.75,-2.05)--(-3.75,2.05)
        node[above] {\(\Phi^*_A\)};
      \node[rotate=90,align=center] at (-4.18,0.0)
        {antifields};
      \node[align=center,text width=4.7cm] at (-0.45,2.05)
        {\(\omega_{\rm BV}
        =\delta\Phi^*_A\wedge\delta\Phi^A\),\\
        \((F,G)=\iota_{X_F}\iota_{X_G}\omega_{\rm BV}\)};

      \draw[very thick,blue!75!black]
        (-3.85,-0.85) .. controls (-2.55,-0.18) and (-1.25,0.05) ..
        (0.05,0.28) .. controls (1.15,0.48) and (2.00,0.88) .. (2.65,1.28);
      \draw[very thick,orange!85!black]
        (-3.85,-1.35) .. controls (-2.40,-0.98) and (-1.05,-0.70) ..
        (0.20,-0.40) .. controls (1.25,-0.15) and (2.05,0.18) .. (2.65,0.62);
      \node[blue!75!black,align=left] at (1.05,1.42)
        {\(\mathcal L_{\Psi_1}\!:\;
        \Phi^*_A=\dfrac{\delta^L\Psi_1}{\delta\Phi^A}\)};
      \node[orange!85!black,align=left] at (0.95,-0.95)
        {\(\mathcal L_{\Psi_2}\!:\;
        \Phi^*_A=\dfrac{\delta^L\Psi_2}{\delta\Phi^A}\)};
      \draw[-{Latex[length=2mm]},thick,gray!80]
        (-2.05,-0.74)--(-2.05,-1.12)
        node[midway,left,black] {\(\Psi_1\to\Psi_2\)};
      \node[align=center,text width=5.6cm] at (-0.45,-2.62)
        {on any \(\mathcal L_\Psi\), \(\omega_{\rm BV}|_{\mathcal L_\Psi}=0\)
        and the integration coordinates are the \(\Phi^A\)};
    \end{scope}

    \begin{scope}[shift={(4.95,0)}]
      \draw[thick,rounded corners=3pt,fill=green!7] (-2.05,1.45) rectangle (2.05,2.35);
      \node[align=center] at (0,1.90)
        {classical master equation\\
        \((S_{\rm BV},S_{\rm BV})=0\)};
      \draw[thick,rounded corners=3pt,fill=purple!7] (-2.05,0.10) rectangle (2.05,1.00);
      \node[align=center] at (0,0.55)
        {BV Hamiltonian vector field\\
        \(Q=(S_{\rm BV},\,\cdot\,)\), \(Q^2=0\)};
      \draw[thick,rounded corners=3pt,fill=red!7] (-2.05,-1.25) rectangle (2.05,-0.35);
      \node[align=center] at (0,-0.80)
        {gauge-fixed action\\
        \(S_\Psi=S_{\rm BV}|_{\mathcal L_\Psi}\)};
      \draw[thick,rounded corners=3pt,fill=yellow!12] (-2.05,-2.60) rectangle (2.05,-1.70);
      \node[align=center] at (0,-2.15)
        {quantum identity\\
        \(\Delta e^{\ii S_{\rm BV}/\hbar}=0\)};
      \draw[-{Latex[length=2mm]},thick] (0,1.45)--(0,1.00);
      \draw[-{Latex[length=2mm]},thick] (0,0.10)--(0,-0.35);
      \draw[-{Latex[length=2mm]},thick] (0,-1.25)--(0,-1.70);
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The BV antibracket is the Poisson bracket associated with the odd
  symplectic form on field--antifield space.  A gauge-fixing fermion
  \(\Psi\) specifies the Lagrangian submanifold
  \(\Phi_A^*=\delta^L\Psi/\delta\Phi^A\); restricting \(S_{\rm BV}\) to this
  submanifold gives the gauge-fixed action.}
  \label{fig:bv-field-antifield-master-map}
\end{figure}

\section{Half-Densities and the BV Integration Measure}
\label{sec:bv-half-densities-measure}

The phrase ``the BV measure'' has a precise finite-dimensional meaning.  Let
\(\mathcal E\) be the finite-dimensional supermanifold of fields and
antifields at a fixed regulator, equipped with the odd symplectic form
\[
  \omega
  =
  \delta\Phi^*_A\,\delta\Phi^A
\]
in Darboux coordinates.  A density on a supermanifold is a section of the
Berezinian line \(\operatorname{Ber}(\mathcal E)\).  A half-density is a
section of \(\operatorname{Ber}(\mathcal E)^{1/2}\).  In a coordinate system
\(z^I=(\Phi^A,\Phi^*_A)\), a half-density is written
\[
  \sigma=f(z)\,|Dz|^{1/2},
\]
where \(f\) transforms by the square root of the Berezinian of the coordinate
change.  This transformation law is exactly what makes the restriction of
\(\sigma\) to a Lagrangian submanifold into an ordinary Berezin density on
that submanifold.

\begin{definition}[Canonical BV operator on half-densities]
\label{def:canonical-bv-half-density-operator}
On an odd symplectic supermanifold \((\mathcal E,\omega)\), the canonical BV
operator on half-densities is the odd second-order operator
\(\Delta_{1/2}\) defined in every Darboux chart by
\begin{equation}
  \Delta_{1/2}\!\left(f\,|D\Phi\,D\Phi^*|^{1/2}\right)
  =
  \left[
  \sum_A
  (-1)^{\epsilon_A}
  \frac{\partial^R}{\partial\Phi^A}
  \frac{\partial^R f}{\partial\Phi^*_A}
  \right]
  |D\Phi\,D\Phi^*|^{1/2}.
  \label{eq:canonical-half-density-bv-operator}
\end{equation}
The expression is invariant under odd-symplectic coordinate changes and
\(\Delta_{1/2}^2=0\).  This is the finite-dimensional Khudaverdian
semidensity operator; the displayed Darboux formula is the local expression
used by perturbative regulators.
\end{definition}

\begin{definition}[Normal BV half-density data]
\label{def:normal-bv-half-density-data}
A finite-dimensional quantum BV regulator contains, in addition to
\((\mathcal E,\omega)\), an even nowhere-vanishing reference half-density
\(\sigma_0\).  It is called normal when
\[
  \Delta_{1/2}\sigma_0=0 .
\]
For normal half-density data, the BV Laplacian on functions is the divergence operator
\begin{equation}
  \Delta_{\sigma_0}F
  :=
  \sigma_0^{-1}\Delta_{1/2}(F\sigma_0).
  \label{eq:function-bv-laplacian-from-half-density}
\end{equation}
In a Darboux chart in which \(\sigma_0=|D\Phi\,D\Phi^*|^{1/2}\), this becomes
\begin{equation}
  \Delta_{\sigma_0}
  =
  \sum_A
  (-1)^{\epsilon_A}
  \frac{\partial^R}{\partial\Phi^A}
  \frac{\partial^R}{\partial\Phi^*_A}.
  \label{eq:bv-laplacian}
\end{equation}
The square \(\sigma_0^2\) is the corresponding Berezinian density.  If a
chosen regulator supplies a reference half-density with
\(\Delta_{1/2}\sigma_0\ne0\), its logarithmic defect is an additional local
measure term; the symbol \(\Delta_{\rm BV}\) in this chapter denotes the
function operator \(\Delta_{\sigma_0}\) only after normal half-density data
have been chosen.  In Schwarz's geometric terminology, the pair consisting of
the odd symplectic structure and these compatible volume/half-density data forms
the finite-dimensional measure structure behind BV quantization.
\end{definition}

\begin{proposition}[Restriction to Lagrangians and BV Stokes]
\label{prop:bv-half-density-restriction-stokes}
Let \(L\subset\mathcal E\) be a Lagrangian submanifold with embedding
\(\iota_L\).  Every half-density \(\sigma\) on \(\mathcal E\) determines a
Berezin density \(\iota_L^*\sigma\) on \(L\).  In a Darboux chart adapted to
\(L\), so that \(L\) is given by \(\Phi^*_A=0\),
\[
  \iota_L^*\!\left(f(\Phi,\Phi^*)\,|D\Phi\,D\Phi^*|^{1/2}\right)
  =
  f(\Phi,0)\,|D\Phi|.
\]
If \(\eta\) is a compactly supported half-density, or if the chosen
integration cycle has boundary behavior that kills total derivatives, then
\begin{equation}
  \int_L \iota_L^*(\Delta_{1/2}\eta)=0 .
  \label{eq:bv-stokes-half-density}
\end{equation}
\end{proposition}

\begin{proof}
In an adapted Darboux chart the stated restriction formula gives a Berezin
density on the coordinates of \(L\).  Under a change of adapted Darboux
coordinates, the half-density coefficient transforms by the square root of
the Berezinian on the full odd symplectic space, and the Lagrangian condition
identifies this factor with the Berezinian transformation of the induced
coordinates on \(L\).  Hence the local formula glues to an intrinsic density.

For the Stokes statement, write
\(\eta=h(\Phi,\Phi^*)|D\Phi\,D\Phi^*|^{1/2}\) in an adapted chart.  Restricting
\(\Delta_{1/2}\eta\) to \(L\) gives
\[
  \sum_A
  (-1)^{\epsilon_A}
  \frac{\partial^R}{\partial\Phi^A}
  \left.
  \frac{\partial^R h}{\partial\Phi^*_A}
  \right|_{\Phi^*=0}
  |D\Phi|,
\]
a total divergence on \(L\).  Its integral vanishes under the compact-support
or boundary hypothesis.  A partition of unity gives the general finite
dimensional statement.
\end{proof}

\section{The Quantum Master Equation}

The classical master equation controls the algebra of gauge symmetry.  The
quantum functional integral also depends on the half-density data just
defined.  In field theory the operator \(\Delta_{\sigma_0}\) requires a
regulator because the two functional derivatives act at the same point.  A
quantum BV problem therefore specifies the regularized odd symplectic space,
the normal reference half-density \(\sigma_0\), and the action.

\begin{definition}[Quantum master equation]
For Lorentzian weight \(\exp(\ii S/\hbar)\), the BV integrand is the
half-density
\[
  \sigma_S
  :=
  \exp(\ii S/\hbar)\,\sigma_0 .
\]
The quantum master equation is the intrinsic condition
\begin{equation}
  \Delta_{1/2}\sigma_S=0 .
  \label{eq:qme-half-density-form}
\end{equation}
When \(\sigma_0\) is normal, \eqref{eq:qme-half-density-form} is equivalent to
\begin{equation}
  \frac12(S,S)-\ii\hbar\,\Delta_{\sigma_0}S=0.
  \label{eq:quantum-master-equation}
\end{equation}
Equivalently,
\[
  \Delta_{\sigma_0}\exp(\ii S/\hbar)=0.
\]
\end{definition}

\begin{proposition}[Component form of the quantum master equation]
\label{prop:qme-component-form}
For normal BV half-density data and an even action \(S\),
\[
  \Delta_{\sigma_0}\exp(\ii S/\hbar)=0
\]
is equivalent to
\[
  \frac12(S,S)-\ii\hbar\,\Delta_{\sigma_0}S=0 .
\]
\end{proposition}

\begin{proof}
In Darboux coordinates with normal reference half-density, the BV Laplacian
is an odd second-order operator.  For homogeneous functions \(F,G\), its
failure to be a derivation is the antibracket:
\begin{equation}
  \Delta_{\sigma_0}(FG)
  =
  (\Delta_{\sigma_0}F)G
  +
  (-1)^{|F|}F\Delta_{\sigma_0}G
  +
  (-1)^{|F|}(F,G).
  \label{eq:bv-laplacian-second-order-identity}
\end{equation}
This identity follows by applying
\eqref{eq:bv-laplacian} to the product \(FG\) and collecting the two
terms in which one derivative hits each factor.  Since \(S\) is even,
induction using \eqref{eq:bv-laplacian-second-order-identity} gives
\begin{equation}
  \Delta_{\sigma_0}S^n
  =
  nS^{n-1}\Delta_{\sigma_0}S
  +
  \frac{n(n-1)}{2}S^{n-2}(S,S).
  \label{eq:bv-laplacian-power-even-action}
\end{equation}
Expanding the exponential and summing the power series yields
\[
  \Delta_{\sigma_0}\ee^{\ii S/\hbar}
  =
  \left[
    \frac{\ii}{\hbar}\Delta_{\sigma_0}S
    -
    \frac{1}{2\hbar^2}(S,S)
  \right]\ee^{\ii S/\hbar}.
\]
The exponential factor is invertible as a formal power series, so the
vanishing of the left hand side is equivalent to the displayed component
equation.
\end{proof}

The half-density form is the coordinate-free statement.  Gauge independence is
then a version of Proposition~\ref{prop:bv-half-density-restriction-stokes}:
changing the gauge-fixing Lagrangian submanifold changes the integral by the
integral of a \(\Delta_{1/2}\)-exact half-density, provided no boundary
contribution is present.

\begin{definition}[Gauge-fixed Wilsonian BV consistency]
\label{def:gauge-fixed-wilsonian-bv-consistency}
At cutoff \(\Lambda\), a gauge-fixed Wilsonian path integral is BV-consistent
when it is obtained by restricting a BV half-density
\[
  \sigma_\Lambda
  =
  \exp(\ii S_\Lambda/\hbar)\sigma_{0,\Lambda}
\]
to a gauge-fixing Lagrangian submanifold
\(\mathcal L_{\Psi,\Lambda}\subset\mathcal E_\Lambda\), with
\[
  \Delta_{1/2,\Lambda}\sigma_\Lambda=0 .
\]
Equivalently, for normal \(\sigma_{0,\Lambda}\), the cutoff action satisfies
\[
  \frac12(S_\Lambda,S_\Lambda)_\Lambda
  -\ii\hbar\,\Delta_{\sigma_{0,\Lambda}}S_\Lambda=0 .
\]
The gauge-fixed source functional is the restriction
\[
  Z_{\Lambda,\Psi}[J]
  =
  \int_{\mathcal L_{\Psi,\Lambda}}
  \iota_{\Psi,\Lambda}^*
  \left[
    \exp\left\{\frac{\ii}{\hbar}
    \left(S_\Lambda+\langle J,\Phi\rangle\right)\right\}
    \sigma_{0,\Lambda}
  \right].
\]
\end{definition}

The definition identifies the master equation as the gauge-consistency
condition for the gauge-fixed Wilsonian integral.  The gauge condition
\(\Psi\) selects a Lagrangian submanifold for computation; the equation lives
on the field-antifield space before this restriction.  Differentiating the
restricted integral and applying BV Stokes gives the cutoff
Slavnov--Taylor identities.  Changing \(\Psi\) changes the representative of
the same BV half-density integral when the boundary term in
\eqref{eq:bv-stokes-half-density} vanishes.  A failure of the displayed
master equation is a ghost-number-one anomaly functional, and
\eqref{eq:qme-defect-consistency} is its Wess--Zumino consistency condition.

If the regulated theory fails to satisfy \eqref{eq:quantum-master-equation},
the failure has ghost number \(+1\).  To first order in \(\hbar\), write
\begin{equation}
  \mathcal M(S)
  :=
  \frac12(S,S)-\ii\hbar\,\Delta_{\sigma_0}S
  =
  \hbar\,\mathcal A+O(\hbar^2).
  \label{eq:qme-anomaly-defect}
\end{equation}
The BV Jacobi identity, the second-order identity for \(\Delta_{\sigma_0}\),
and \(\Delta_{\sigma_0}^2=0\) give the quantum consistency identity
\begin{equation}
  (S,\mathcal M(S))-\ii\hbar\,\Delta_{\sigma_0}\mathcal M(S)=0.
  \label{eq:qme-defect-consistency}
\end{equation}
Indeed, for even \(S\), \((S,\frac12(S,S))=0\) and
\(\Delta_{\sigma_0}(S,S)=2(\Delta_{\sigma_0}S,S)\), so the two order-\(\hbar\)
terms in \eqref{eq:qme-defect-consistency} cancel.  Substituting
\eqref{eq:qme-anomaly-defect} and \(S=S_0+O(\hbar)\) into
\eqref{eq:qme-defect-consistency} gives
\[
  0
  =
  \hbar\,(S_0,\mathcal A)+O(\hbar^2).
\]
Equivalently, with \(s_{\rm BV}:=(\cdot,S_0)\),
\[
  s_{\rm BV}\mathcal A=(\mathcal A,S_0)=0 .
\]
Changing the local counterterm by \(\hbar B\) changes the anomaly by
\begin{equation}
  \mathcal A\longmapsto \mathcal A+s_{\rm BV}B
  =
  \mathcal A+(B,S_0) .
  \label{eq:qme-anomaly-counterterm-shift}
\end{equation}
Thus the obstruction class lies in ghost-number-one local BV cohomology.  This
is the antifield version of the anomaly classification by
\(H^{1,D}(s\mid d)\).  In the chiral gauge example of
Chapter~\ref{ch:volume-ii-20}, the representative is the
descent cocycle whose bottom component gives the consistent anomaly
\eqref{eq:consistent-descent-anomaly-representative}; adding a local
counterterm changes the representative by an \(s_{\rm BV}\)-exact cocycle,
matching \eqref{eq:qme-anomaly-counterterm-shift}.

\section{The 1PI Master Equation}

The antifields are also the natural sources for nonlinear BRST variations in
the 1PI formalism.  With a gauge-fixing Lagrangian submanifold chosen for the
fields, let \(\iota_\Psi:\mathcal L_\Psi\hookrightarrow\mathcal E\) be its
embedding and let \(\Phi^*_{\rm ext}\) denote the antifield coordinates left
as external sources after the gauge-fixing canonical transformation.  Define
\[
  Z[J,\Phi^*_{\rm ext}]
  =
  \int_{\mathcal L_\Psi}
  \iota_\Psi^*
  \left[
  \exp\left\{
  \frac{\ii}{\hbar}
  \left(S_{\rm BV}[\Phi,\Phi^*_{\rm ext}]+\int J_A\Phi^A\right)
  \right\}
  \sigma_0
  \right].
\]
In an adapted finite-regulator chart this is the compact formula with
\[
  D_{\sigma_0,\Psi}\Phi:=\iota_\Psi^*\sigma_0
\]
in place of the undefined symbol \([D\Phi]\).  The external antifields are not
integrated; they record the response to BRST variations.  Let
\(W[J,\Phi^*_{\rm ext}]\) be the connected generator and let
\[
  \varphi^A=\frac{\delta W}{\delta J_A},
  \qquad
  \Gamma[\varphi,\Phi^*_{\rm ext}]
  =
  W[J,\Phi^*_{\rm ext}]-\int J_A\varphi^A
\]
be the 1PI effective action in the Lorentzian convention used earlier.

Assume that the regularized half-density integrand obeys the quantum master
equation at \(J=0\) and that the Legendre transform is well defined on the
field directions under consideration.  The Ward identity obtained by applying
\(\Delta_{1/2}\) to the source-dependent integrand and then using BV Stokes
gives
\[
  \int
  \frac{\delta^R\Gamma}{\delta\varphi^A}
  \frac{\delta^L\Gamma}{\delta\Phi^*_{\rm ext,A}}
  \dd^D x
  =
  0 .
\]
Equivalently,
\begin{equation}
  \frac12(\Gamma,\Gamma)=0,
  \label{eq:1pi-master-equation}
\end{equation}
where the bracket now pairs the mean field \(\varphi^A\) with the antifield
\(\Phi^*_{\rm ext,A}\).  In the Yang--Mills variables of the preceding chapter,
\eqref{eq:1pi-master-equation} is the Slavnov--Taylor identity
\eqref{eq:slavnov-taylor-identity}, with \(A^*\) and \(c^*\) replacing the
special sources \(K\) and \(L\), and with the nonminimal term producing
\(B\,\delta\Gamma/\delta\bar c\).
The all-order restoration theorem
\ref{thm:all-order-slavnov-taylor-restoration} is therefore the gauge-fixed
Yang--Mills form of the perturbative BV statement: when the local
ghost-number-one obstruction class vanishes, counterterms may be chosen so that
the renormalized 1PI functional satisfies \eqref{eq:1pi-master-equation} as a
formal power series in \(\hbar\).  The ghost-number-zero BV cohomology records
the remaining finite coordinate freedom.

\section{Wilsonian BV Pushforward}

The Wilsonian effective action is obtained by integrating out degrees of
freedom above a scale.  BV gives the condition under which this operation is
compatible with gauge symmetry.

The input is more than a momentum cutoff.  A gauge-compatible Wilsonian step
contains a finite-dimensional or regularized BV phase space
\((\mathcal E_\Lambda,\omega_\Lambda)\), a normal half-density
\(\sigma_{0,\Lambda}\), a solution \(S_\Lambda\) of the quantum master
equation, a split into low and high BV variables, and a Lagrangian cycle for
the high variables.  Equivalently, in a lattice formulation it contains a
finite lattice gauge integral and a gauge-invariant blocking map for
observables or actions.  These data are what make the phrase "integrating out
gauge-field modes" a defined operation.

For a gauge-fixed Wilsonian path integral, the flow between cutoffs preserves
Definition~\ref{def:gauge-fixed-wilsonian-bv-consistency}.  In BV terms this
means that the high-variable integration is a BV pushforward of the cutoff
half-density.  The resulting low-cutoff half-density again satisfies the
quantum master equation, and every gauge-fixed representative obtained from
it satisfies the corresponding cutoff Slavnov--Taylor identities.

Consider first a finite-dimensional odd symplectic vector space split as
\[
  \mathcal E=\mathcal E_{\rm low}\oplus\mathcal E_{\rm high},
  \qquad
  \Delta_{1/2,\mathcal E}
  =
  \Delta_{1/2,{\rm low}}+\Delta_{1/2,{\rm high}}.
\]
Assume the normal half-density factorizes compatibly as
\[
  \sigma_0=\sigma_{0,{\rm low}}\otimes\sigma_{0,{\rm high}} .
\]
Let \(\mathcal L_{\rm high}\subset\mathcal E_{\rm high}\) be a Lagrangian
subspace used to integrate out the high variables.  Define the effective
low-energy half-density by
\begin{equation}
  \exp(\ii S_{\rm eff}/\hbar)\,\sigma_{0,{\rm low}}
  =
  \int_{\mathcal L_{\rm high}}
  \iota_{\rm high}^*
  \left[
    \exp(\ii S/\hbar)\,\sigma_0
  \right].
  \label{eq:bv-wilsonian-pushforward}
\end{equation}

\begin{proposition}[BV pushforward]
Assume
\[
  \Delta_{1/2,\mathcal E}
  \left(\exp(\ii S/\hbar)\sigma_0\right)=0
\]
and assume the integral over \(\mathcal L_{\rm high}\) has no boundary
contribution for \(\Delta_{1/2,\rm high}\).  Then
\[
  \Delta_{1/2,{\rm low}}
  \left(\exp(\ii S_{\rm eff}/\hbar)\sigma_{0,{\rm low}}\right)=0 .
\]
\end{proposition}

\begin{proof}
Using the compatible split
\(\Delta_{1/2,\mathcal E}=\Delta_{1/2,{\rm low}}+\Delta_{1/2,{\rm high}}\),
\[
\begin{aligned}
  \Delta_{1/2,{\rm low}}
  \left(\exp(\ii S_{\rm eff}/\hbar)\sigma_{0,{\rm low}}\right)
  &=
  \int_{\mathcal L_{\rm high}}
  \iota_{\rm high}^*
  \left[
    \Delta_{1/2,{\rm low}}
    \left(\exp(\ii S/\hbar)\sigma_0\right)
  \right]
\\
  &=
  -\int_{\mathcal L_{\rm high}}
  \iota_{\rm high}^*
  \left[
    \Delta_{1/2,{\rm high}}
    \left(\exp(\ii S/\hbar)\sigma_0\right)
  \right]
  =
  0 .
\end{aligned}
\]
The last equality is Proposition~\ref{prop:bv-half-density-restriction-stokes}
applied to the high Lagrangian integration cycle.
\end{proof}

This proposition is the regulated mathematical core of the Wilsonian BV
identity.  In a field theory with cutoff \(\Lambda\), the Wilsonian action
\(S_\Lambda\) is a scale-dependent local or quasi-local representative of the
BV pushforward.  Its master equation has the form
\[
  \frac12(S_\Lambda,S_\Lambda)_\Lambda
  -
  \ii\hbar\,\Delta_\Lambda S_\Lambda
  =
  0,
\]
where the bracket and Laplacian include the cutoff covariance and the
remaining low-mode half-density.  Exact RG flow is compatible with gauge
symmetry when it preserves this equation.  Changes of cutoff scheme are then
BV canonical transformations together with possible local counterterm changes,
with the gauge symmetry encoded by the transformed master action.

The lattice route gives a second gauge-compatible Wilsonian construction.  A
finite blocking map sends link variables on a fine lattice to link variables
or gauge-invariant observables on a coarser lattice, and the pushforward of
the compact Haar integral defines the coarse action or coarse observable
law.  The continuum Wilsonian statement then concerns a scaling limit of these
finite lattice pushforwards.  A comparison between the lattice and BV
descriptions must identify the gauge-invariant observables, the local BV
cohomology classes, and the limiting topology for correlation functions or
effective-action coordinates.

\section{Infinitesimal Exact RG in BV Form}
\label{sec:bv-infinitesimal-exact-rg}

The finite pushforward above is the primary definition.  A differential
renormalization-group equation is obtained from a differentiable family of
finite pushforwards or from an infinitesimal change of regulator.  The
following finite-regulator statement records exactly what must be preserved.

\begin{definition}[BV-compatible infinitesimal RG step]
\label{def:bv-compatible-infinitesimal-rg-step}
Let \((\mathcal E,\omega,\sigma_0)\) be a finite-dimensional odd symplectic
space with normal half-density data and let
\[
  \rho_t=\exp(\ii S_t/\hbar)\sigma_0
\]
be an even family of half-densities.  An infinitesimal RG step is
BV-compatible if there is an odd ghost-number \(-1\) function \(R_t\) such
that
\begin{equation}
  \partial_t\rho_t
  =
  \Delta_{1/2}(R_t\rho_t).
  \label{eq:bv-infinitesimal-rg-density}
\end{equation}
The function \(R_t\) is part of the scheme data.  It encodes the infinitesimal
change of covariance, splitting, or gauge-fixing cycle after the regulator has
made these words finite.
\end{definition}

\begin{theorem}[Infinitesimal BV RG preserves the QME]
\label{thm:infinitesimal-bv-rg-preserves-qme}
Assume \(\Delta_{1/2}\rho_{t_0}=0\) at one value \(t_0\), and assume
\eqref{eq:bv-infinitesimal-rg-density} holds on an interval with the same
normal half-density data.  Then
\[
  \Delta_{1/2}\rho_t=0
\]
for all \(t\) in that interval.  In action coordinates, whenever \(S_t\) is
defined by \(\rho_t=\exp(\ii S_t/\hbar)\sigma_0\), the evolution is
\begin{equation}
  \partial_t S_t
  =
  (S_t,R_t)-\ii\hbar\,\Delta_{\sigma_0}R_t ,
  \label{eq:bv-infinitesimal-rg-action}
\end{equation}
provided \(S_t\) satisfies the quantum master equation.
\end{theorem}

\begin{proof}
Apply \(\Delta_{1/2}\) to
\eqref{eq:bv-infinitesimal-rg-density}.  Since the canonical operator on
half-densities obeys \(\Delta_{1/2}^2=0\),
\[
  \partial_t(\Delta_{1/2}\rho_t)
  =
  \Delta_{1/2}^2(R_t\rho_t)=0 .
\]
Thus \(\Delta_{1/2}\rho_t\) is constant in \(t\), and the assumed vanishing at
\(t_0\) gives the result.

For the action form, write \(\rho_t=\ee^{\ii S_t/\hbar}\sigma_0\) and use the
function Laplacian induced by the normal half-density.  Since \(R_t\) is odd,
the second-order identity gives
\[
  \Delta_{\sigma_0}(R_t\ee^{\ii S_t/\hbar})
  =
  (\Delta_{\sigma_0}R_t)\ee^{\ii S_t/\hbar}
  -
  (R_t,\ee^{\ii S_t/\hbar}).
  \]
The bracket term is
\[
  (R_t,\ee^{\ii S_t/\hbar})
  =
  \frac{\ii}{\hbar}(R_t,S_t)\ee^{\ii S_t/\hbar}
  =
  -\frac{\ii}{\hbar}(S_t,R_t)\ee^{\ii S_t/\hbar},
\]
because \(S_t\) is even.  Dividing
\[
  \partial_t\ee^{\ii S_t/\hbar}
  =
  \Delta_{\sigma_0}(R_t\ee^{\ii S_t/\hbar})
\]
by \(\ee^{\ii S_t/\hbar}\) and multiplying by \(\hbar/\ii\) gives
\eqref{eq:bv-infinitesimal-rg-action}.
\end{proof}

The equation is a quantum canonical flow, not an arbitrary beta-function
ansatz.  It is the differential version of BV pushforward: the right hand side
is \(\Delta_{1/2}\)-exact on half-densities, so it cannot create a master
equation defect.  A cutoff flow proposed for a gauge theory is therefore
gauge-compatible only after its regulator supplies \(R_t\), the odd symplectic
structure, the half-density data, and the boundary condition making the
finite pushforward or the differential equation legitimate.

\section{Counterterms, Anomalies, and Physical Operators}

Let \(S_{\rm BV}\) be a classical solution of the master equation and let
\[
  s_{\rm BV}=(\cdot,S_{\rm BV}).
\]
A first-order local deformation
\[
  S_{\rm BV}\longmapsto S_{\rm BV}+\epsilon C
\]
preserves the classical master equation to order \(\epsilon\) precisely when
\[
  s_{\rm BV}C=0,
  \qquad
  \operatorname{gh}(C)=0.
\]
If \(C=s_{\rm BV}B\), then the deformation is generated by a BV canonical
transformation and does not change the physical theory.  Thus invariant
counterterms are classified by ghost-number-zero BV cohomology of local
functionals, modulo total derivatives and canonical redefinitions.

An anomaly candidate is a ghost-number-one functional \(\mathcal A\) satisfying
\[
  s_{\rm BV}\mathcal A=0.
\]
It is removable by a local counterterm exactly when
\[
  \mathcal A=s_{\rm BV}B .
\]
This gives the precise cohomological meaning of anomaly cancellation in the
perturbative BV framework: the obstruction class must vanish in the relevant
local BV cohomology.

Physical local operators are treated in the same language.  An insertion
\(\mathcal O\) is compatible with the gauge symmetry when
\[
  s_{\rm BV}\mathcal O=0,
  \qquad
  \operatorname{gh}(\mathcal O)=0,
\]
and two insertions differing by \(s_{\rm BV}\mathcal X\) define the same
cohomology class.  After gauge fixing, this statement reduces to the BRST
cohomology of local observables described in the preceding chapter.  Before
gauge fixing, it is expressed without choosing a particular gauge slice.

\section{BV--BRST Equivalence for Gauge-Fixed Yang--Mills}
\label{sec:bv-brst-equivalence-ym}

The gauge-fixed BRST complex used in the preceding chapter is obtained from
the BV complex by two homological operations: a BV canonical transformation
generated by the gauge-fixing fermion, and deletion of contractible
nonminimal pairs.  The next theorem states the equivalence in the form needed
for perturbative Yang--Mills with matter.

\begin{theorem}[Local BV--BRST comparison for Yang--Mills]
\label{thm:bv-brst-local-comparison-ym}
Let \(G\) be compact and reductive, and let the classical Yang--Mills--matter
action use the trace and coupling convention of
Chapter~\ref{ch:vol2-classical-yang-mills-matter},
\[
  S_{\rm YM}=\frac1{4g_{\rm YM}^2}\int\operatorname{tr}(F_{\mu\nu}F^{\mu\nu})
  +S_{\rm matter}.
\]
Assume the perturbative local jet algebra is taken near a background for
which the Koszul--Tate differential is acyclic in positive antifield number
for the fields under consideration, and assume boundary conditions that make
local functionals modulo total derivatives the correct cochain space.  Then:
\begin{enumerate}
\item The canonical BV transformation generated by a gauge-fixing fermion
\(\Psi\) preserves the antibracket and the classical master equation.
\item Restriction to the Lagrangian submanifold
\(\Phi^*=\delta^L\Psi/\delta\Phi\) gives the gauge-fixed BRST differential on
fields, with the nonminimal pair \(s\bar c=B\), \(sB=0\).
\item The inclusion of antifield-independent representatives induces an
isomorphism, for nonnegative ghost number, between local BV cohomology modulo
\(\dd\) and gauge-fixed BRST cohomology modulo \(\dd\), after quotienting the
contractible pair \((\bar c,B)\):
\[
  H^{g,D}(s_{\rm BV}\mid\dd)
  \cong
  H^{g,D}(s_{\Psi}\mid\dd),
  \qquad g\ge0 .
\]
The isomorphism identifies counterterm classes, anomaly classes, and
gauge-invariant local operator classes.
\end{enumerate}
\end{theorem}

\begin{proof}
The finite-regulator BV phase space has Darboux coordinates
\((\Phi^A,\Phi^*_A)\).  The gauge-fixing fermion generates the canonical
change of variables
\[
  \Phi^A_{\Psi}=\Phi^A,\qquad
  \Phi^*_{\Psi,A}=\Phi^*_A-\frac{\delta^L\Psi}{\delta\Phi^A}.
\]
It is canonical because
\[
  \delta\Phi^*_{\Psi,A}\,\delta\Phi^A_{\Psi}
  =
  \delta\Phi^*_A\,\delta\Phi^A
  -
  \delta\!\left(\frac{\delta^L\Psi}{\delta\Phi^A}\right)\delta\Phi^A ,
\]
and the second term is the odd-symplectic differential of the exact one-form
generated by \(\Psi\); its antisymmetrized second variation vanishes.  Hence
the antibracket is unchanged, and \((S_{\rm BV},S_{\rm BV})=0\) is equivalent
to the same equation in the transformed coordinates.  Setting
\(\Phi^*_{\Psi}=0\) is precisely the Lagrangian restriction
\(\Phi^*=\delta^L\Psi/\delta\Phi\).

On antifield-independent functions of the fields, the Hamiltonian vector
field \((\,\cdot\,,S_{\rm BV})\) restricted to this Lagrangian is the
gauge-fixed BRST variation.  For Yang--Mills the displayed minimal action
gives \(sA=Dc\), \(sc=-\frac12[c,c]\), and the nonminimal term gives
\(s\bar c=B\), \(sB=0\).  The proof of the classical master equation above
therefore gives \(s_\Psi^2=0\) before eliminating \(B\).

It remains to compare cohomology.  Filter the BV local functional complex by
antifield number.  The antifield-number \(-1\) piece of \(s_{\rm BV}\) is the
Koszul--Tate differential \(\delta_{\rm KT}\), with
\(\delta_{\rm KT}\Phi^*_i=E_i\) and
\(\delta_{\rm KT}c^*_\alpha=R^i{}_\alpha\Phi^*_i\).  By the stated regularity
hypothesis, \(\delta_{\rm KT}\) has no cohomology in positive antifield
number.  Therefore the spectral sequence of the antifield-number filtration
collapses to representatives of antifield number zero for \(g\ge0\), modulo
the gauge variation induced by the longitudinal BRST part.

The nonminimal pair is contractible.  Let \(N_{\rm nm}\) count the number of
\(\bar c\), \(B\), and their derivatives in a local monomial, and let
\[
  h_{\rm nm}
  =
  \sum_{|\alpha|\ge0}
  \partial_\alpha\bar c^a
  \frac{\partial^L}{\partial(\partial_\alpha B^a)} .
\]
On the subalgebra generated by the nonminimal pair and its derivatives,
\[
  s h_{\rm nm}+h_{\rm nm}s=N_{\rm nm}.
\]
If a cocycle has \(N_{\rm nm}>0\), division by this positive integer writes
that component as an \(s\)-exact term.  Hence the nonminimal sector contributes
no cohomology.  Combining this doublet contraction with the
Koszul--Tate acyclicity gives the displayed isomorphism.
\end{proof}

\section{Local Cohomology Ledger for Yang--Mills With Matter}
\label{sec:bv-local-cohomology-ym}

The comparison theorem turns the abstract statement "a BV class" into a
computable local object.  In four-dimensional power-counting-renormalizable
Yang--Mills with scalars and spinors in finite-dimensional unitary
representations, the local classes used by perturbation theory have the
following concrete representatives.

\begin{proposition}[Renormalizable ghost-number-zero representatives]
\label{prop:bv-renormalizable-ghost-zero-cohomology}
Under the hypotheses of Theorem~\ref{thm:bv-brst-local-comparison-ym}, every
ghost-number-zero counterterm density of engineering dimension at most four is
cohomologous, modulo total derivatives and field redefinitions, to a gauge
invariant density built from covariant fields:
\[
  F_{\mu\nu},\qquad
  D_\mu,\qquad
  \psi,\bar\psi,\qquad
  \phi ,
\]
with all gauge indices contracted by invariant tensors of the representation.
For the pure gauge sector this gives
\[
  \frac1{4}\operatorname{tr}(F_{\mu\nu}F^{\mu\nu}),
  \qquad
  \frac1{4}\operatorname{tr}(F_{\mu\nu}\widetilde F^{\mu\nu})
\]
in the monograph trace convention
\(\operatorname{tr}_{\square}(t^at^b)=\delta^{ab}\).  With matter, the same
list includes covariant kinetic terms, mass terms allowed by the global
symmetries, Yukawa couplings, and scalar potential terms whose gauge indices
are contracted invariantly.  Antifield-dependent ghost-number-zero
representatives encode field redefinitions, coupling-coordinate changes, and
global-symmetry currents rather than new gauge-invariant interaction
densities.
\end{proposition}

\begin{proof}
By Theorem~\ref{thm:bv-brst-local-comparison-ym}, a class may be represented
in the gauge-fixed BRST complex without the nonminimal pair.  The doublet
argument removes \((\bar c,B)\), and the Koszul--Tate part removes
antifield-dependent pieces except for the known field-redefinition and current
classes.  A remaining ghost-number-zero representative is therefore a local
polynomial in \(A,\psi,\bar\psi,\phi\) and their derivatives, annihilated by
the longitudinal gauge BRST differential
\[
  s_{\rm gauge}A=Dc,\qquad
  s_{\rm gauge}X=\ii c^at_R^aX
\]
for a matter field \(X\) in representation \(R\).  This condition is precisely
infinitesimal gauge invariance of the corresponding local density.  The
replacement of ordinary derivatives by covariant derivatives follows by
commuting derivatives until all noncovariant gauge variations are paired into
curvatures \(F_{\mu\nu}\) or covariant derivatives.  Power counting in four
dimensions leaves exactly the displayed pure gauge densities and the covariant
matter terms of dimension at most four.  Total derivatives are zero in the
relative cohomology \(H^{0,4}(s\mid\dd)\), and BRST-exact terms are generated
by BV canonical transformations, namely field and coordinate redefinitions.
\end{proof}

\begin{quotedtheorem}[Ghost-number-one anomaly calculation]
\label{qthm:bv-ghost-one-anomaly-calculation}
For a four-dimensional chiral Yang--Mills theory in the same perturbative
local setting, a perturbative local gauge-anomaly candidate is represented in
\[
  H^{1,4}(s_{\rm BV}\mid\dd)
\]
by the descent of the six-form invariant polynomial
\[
  I_6
  =
  \frac{\ii}{24\pi^2}
  \operatorname{Tr}_R(\mathsf F^3)
\]
in the anti-Hermitian background-connection convention of
Chapter~\ref{ch:volume-ii-20}.  Its component coefficient is proportional to
the symmetric invariant tensor
\[
  d_R^{abc}:=\operatorname{Tr}_R(t^{(a}t^bt^{c)}).
\]
If the sum of \(d_R^{abc}\) over all left-handed Weyl fermions vanishes for
every triple of gauge generators, then the perturbative local gauge anomaly
class vanishes.  If it does not vanish, no local counterterm restores the
quantum master equation in this perturbative local cohomology.
\end{quotedtheorem}

Equation~\eqref{eq:qme-anomaly-defect} and the consistency identity
\eqref{eq:qme-defect-consistency} show that the first nonzero QME defect is a
ghost-number-one \(s_{\rm BV}\)-cocycle modulo total derivatives.  The
BV--BRST comparison theorem identifies this group with the gauge-fixed
relative BRST cohomology.  The descent construction of
Chapter~\ref{ch:volume-ii-20} starts from the closed invariant polynomial
\(\operatorname{Tr}_R(\mathsf F^3)\), writes it locally as
\(\dd I_5^{(0)}\), and then applies \(s\) to obtain
\[
  s I_5^{(0)}+\dd I_4^{(1)}=0 .
\]
The integrated four-form \(I_4^{(1)}\) is the consistent anomaly
representative.  Its cubic group factor is
\(\operatorname{Tr}_R(t^a\{t^b,t^c\})\), equivalently the symmetric tensor
\(d_R^{abc}\) after symmetrization over external gauge fields.  Adding a local
counterterm changes \(I_4^{(1)}\) by an \(s\)-exact term plus a total
derivative, as in \eqref{eq:qme-anomaly-counterterm-shift}.  Therefore the
class vanishes exactly when the total symmetric invariant coefficient
vanishes in the chosen representation content.

The word "anomaly" in this proposition refers to the perturbative local
obstruction detected by \(H^{1,4}(s_{\rm BV}\mid\dd)\).  Global gauge
anomalies, determinant-line topology, and finite background-anomaly phases are
different objects; they are treated in
Chapter~\ref{ch:volume-ii-21} and in the later chapters on
invertible field theories and eta invariants.

\section{Local Quantum Observables as a BV Factorization Algebra}
\label{sec:bv-factorization-algebra-boundary}

The BV effective-action formalism has a local-observable counterpart.  This is
the perturbative factorization-algebra language.  It is useful here because it
states locality at the level of observables rather than only at the level of a
single action functional.

\begin{definition}[Prefactorization algebra of BV observables]
\label{def:bv-prefactorization-algebra}
Let \(M\) be spacetime and let \(\operatorname{Open}(M)\) be the category of
open subsets.  A BV prefactorization algebra assigns to each open
\(U\subset M\) a cochain complex
\[
  \operatorname{Obs}_\Lambda(U)
\]
of regulated quantum observables supported in \(U\).  If
\(U_1,\ldots,U_n\subset V\) are pairwise disjoint opens, it assigns a
cochain map
\[
  m_{U_1,\ldots,U_n;V}:
  \operatorname{Obs}_\Lambda(U_1)\otimes\cdots\otimes
  \operatorname{Obs}_\Lambda(U_n)
  \longrightarrow
  \operatorname{Obs}_\Lambda(V),
\]
compatible with permutation of the \(U_i\)'s and with iterated inclusions.
It is a factorization algebra when these maps satisfy descent for Weiss
covers: observables on \(V\) are recovered, up to quasi-isomorphism, from
observables on covers by opens that contain every finite subset of \(V\).
\end{definition}

For a finite BV regulator at scale \(\Lambda\), the cochain differential on
observables is
\begin{equation}
  Q_\Lambda\mathcal O
  :=
  (S_\Lambda,\mathcal O)_\Lambda
  -
  \ii\hbar\,\Delta_\Lambda\mathcal O .
  \label{eq:bv-quantum-observable-differential}
\end{equation}
The bracket and Laplacian are the regulated ones associated with the same
half-density data used in the quantum master equation.  The support
condition in \(\operatorname{Obs}_\Lambda(U)\) means that the kernels of
\(\mathcal O\) are supported in \(U\), with the regulator determining what
class of kernels or jets is allowed.

\begin{proposition}[QME implies the observable differential]
\label{prop:qme-observable-differential}
If \(S_\Lambda\) satisfies
\[
  \frac12(S_\Lambda,S_\Lambda)_\Lambda
  -
  \ii\hbar\,\Delta_\Lambda S_\Lambda=0,
\]
then \(Q_\Lambda^2=0\).  Moreover the disjoint-support product maps in
Definition~\ref{def:bv-prefactorization-algebra} are cochain maps.
\end{proposition}

\begin{proof}
Apply \(Q_\Lambda\) twice:
\[
\begin{aligned}
  Q_\Lambda^2\mathcal O
  &=
  \left(
    \frac12(S_\Lambda,S_\Lambda)_\Lambda
    -
    \ii\hbar\,\Delta_\Lambda S_\Lambda,
    \mathcal O
  \right)_\Lambda .
\end{aligned}
\]
The equality follows from the odd Jacobi identity, the second-order identity
for \(\Delta_\Lambda\), and \(\Delta_\Lambda^2=0\); the terms in which
\(\Delta_\Lambda\) hits the bracket cancel the terms in which one bracket
comes from each application of \(Q_\Lambda\).  The displayed quantum master
equation sets the right hand side to zero.

For disjoint supports, the regulated bracket and Laplacian have no
coincident-point contraction between two different factors except the
contractions included in the product map.  Therefore \(Q_\Lambda\) obeys the
graded Leibniz rule with respect to the product of disjoint observables, and
the maps \(m_{U_1,\ldots,U_n;V}\) commute with \(Q_\Lambda\).
\end{proof}

Changing the cutoff from \(\Lambda\) to \(\Lambda'\) gives a cochain map
\[
  W_{\Lambda\Lambda'}:
  \operatorname{Obs}_\Lambda(U)
  \longrightarrow
  \operatorname{Obs}_{\Lambda'}(U)
\]
when it is produced by BV pushforward with the same boundary and locality
data as in Section~\ref{sec:bv-infinitesimal-exact-rg}.  The homotopy-RG
content is the statement
\[
  Q_{\Lambda'}W_{\Lambda\Lambda'}
  =
  W_{\Lambda\Lambda'}Q_\Lambda
\]
and compatibility of \(W_{\Lambda\Lambda'}\) with the disjoint-support
products.  Thus the factorization-algebra form is the observable-side record
of the same BV renormalization problem: local counterterms repair failures of
the cochain and product identities precisely when the relevant local BV
cohomology class is trivial.

\section{Logical Output}

The output of this chapter is the following framework.

\begin{enumerate}
  \item A gauge theory is represented on an enlarged graded field space with
  fields, ghosts, nonminimal variables, and antifields.
  \item The antibracket is the odd symplectic pairing between fields and
  antifields.
  \item The classical master equation constructs a nilpotent differential.
  \item Gauge fixing is restriction to a Lagrangian submanifold determined by
  a gauge-fixing fermion.
  \item The quantum master equation is the regulated half-density statement of
  measure-compatible gauge symmetry.
  \item The 1PI and Wilsonian effective actions in continuum gauge theory are
  defined as BV objects satisfying the corresponding master identities; a
  lattice gauge formulation supplies the alternative finite-cutoff route by
  compact group integrals and gauge-invariant blocking maps.
  \item Counterterms, anomalies, and physical operators are local BV
  cohomology classes at the appropriate ghost numbers.
\end{enumerate}

This is the framework used below whenever gauge-theory effective actions are
discussed beyond elementary Faddeev--Popov perturbation theory.
