\chapter{The Statistical Ising Model and Universality}
\label{ch:volume-ii-ising-universality}
\ConventionsEuclidean

\label{ch:ising-universality}

\section{The Finite Lattice Ensemble}

Fix a spatial dimension \(D\geq 1\).  A finite Ising system is specified by a
finite subset \(\Lambda\subset \mathbb Z^D\), a set \(E(\Lambda)\) of
nearest-neighbor unordered pairs, and spin variables
\[
  s_x\in\{-1,+1\},
  \qquad x\in\Lambda .
\]
The set \(E(\Lambda)\) encodes the boundary convention.  For example, it may
come from free boundary conditions on a box, or from periodic boundary
conditions after identifying opposite faces.  Bulk thermodynamic quantities are
defined by taking \(\Lambda\nearrow\mathbb Z^D\) along a specified family and
then checking which limits are independent of this boundary convention.
For ferromagnetic coupling \(J>0\), external magnetic field \(h\), and inverse
temperature \(\beta=T^{-1}\) in units with Boltzmann's constant equal to one,
the energy of a spin configuration \(s=\{s_x\}_{x\in\Lambda}\) is
\[
  H_\Lambda(s)
  =
  -J\sum_{\langle x y\rangle\in E(\Lambda)}s_xs_y
  -h\sum_{x\in\Lambda}s_x .
\]
The finite-volume partition function and thermal expectation of an observable
\(F:\{-1,+1\}^{\Lambda}\to\mathbb C\) are
\[
  Z_\Lambda(\beta,h)
  =
  \sum_{s_x=\pm1}
  \exp\!\left[-\beta H_\Lambda(s)\right],
  \qquad
  \vev{F}_{\Lambda,\beta,h}
  =
  Z_\Lambda(\beta,h)^{-1}
  \sum_{s_x=\pm1}F(s)\,\ee^{-\beta H_\Lambda(s)} .
\]
This is the finite-lattice row of
Table~\ref{tab:regulator-integration-status-catalog}: at fixed
\(\Lambda\) the object is an ordinary finite probability measure on a finite
set of configurations.  Its use as a regulator for a continuum Euclidean QFT
requires thermodynamic and scaling limits with their own positivity,
locality, and tightness or convergence hypotheses.
At \(h=0\) the model has the global \(\mathbb Z_2\) symmetry
\(s_x\mapsto -s_x\) for all \(x\).
The Wilson--Fisher chapter used the \(\mathbb Z_2\)-even RG chart unless an
odd source was added.  The present chapter includes that odd sector by the
magnetic source: in continuum variables it is the deformation
\(\int h(x)\phi(x)\), and on the lattice it is represented by the
\(-h\sum_x s_x\) term in \(H_\Lambda\).  Setting \(h=0\) gives the symmetric
critical theory, while differentiating with respect to \(h\) at \(h=0\)
inserts the \(\mathbb Z_2\)-odd spin field \(\sigma\).

The same finite probability space can be written in diagonal operator notation.
Let
\[
  V_x=\operatorname{span}\{\ket{+}_x,\ket{-}_x\},
  \qquad
  \Hilb_\Lambda=\bigotimes_{x\in\Lambda}V_x ,
\]
and define commuting spin operators by
\(\hat S_x\ket{s}_\Lambda=s_x\ket{s}_\Lambda\), where
\(\ket{s}_\Lambda=\bigotimes_x\ket{s_x}_x\).  Then
\[
  \hat H_\Lambda
  =
  -J\sum_{\langle x y\rangle}\hat S_x\hat S_y
  -h\sum_x\hat S_x,
  \qquad
  Z_\Lambda(\beta,h)
  =
  \operatorname{Tr}_{\Hilb_\Lambda}\ee^{-\beta\hat H_\Lambda}.
\]
The space \(\Hilb_\Lambda\) in this display is a finite-dimensional
representation of a classical statistical ensemble.  A continuum Hilbert space,
when constructed from the scaling limit of Euclidean correlators, is a further
object obtained after the limit.

Figure~\ref{fig:volume-ii-ising-ensemble} keeps the finite Ising ensemble data
separate from the reconstructed continuum Hilbert space: spins are summed in a
finite-dimensional trace before any scaling limit is taken.

\begin{figure}[H]
  \centering
  \resizebox{0.98\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small] at (1.5,3.2) {classical configurations};
      \foreach \x in {0,1,2,3} {
        \foreach \y in {0,1,2} {
          \fill[black] (\x,\y) circle (1.5pt);
        }
      }
      \foreach \x in {0,1,2} {
        \foreach \y in {0,1,2} {
          \draw[gray!60] (\x,\y)--(\x+1,\y);
        }
      }
      \foreach \x in {0,1,2,3} {
        \foreach \y in {0,1} {
          \draw[gray!60] (\x,\y)--(\x,\y+1);
        }
      }
      \foreach \x/\y/\spin in {0/0/+,1/0/-,2/0/-,3/0/+,0/1/-,1/1/+,2/1/-,3/1/-,0/2/+,1/2/+,2/2/-,3/2/+} {
        \node[blue!70!black] at (\x,\y+0.28) {\(\spin\)};
      }
      \node[align=center] at (1.5,-0.65)
        {\(s_x=\pm1\), \(x\in\Lambda\subset\mathbb Z^D\)};
    \end{scope}

    \begin{scope}[shift={(5.0,0)}]
      \node[font=\small] at (0,3.2) {Boltzmann weight};
      \node[draw=black, rounded corners, align=center, inner sep=7pt] at (0,1.55)
        {\(\displaystyle
        \mathbb P_{\beta,h}(s)
        =
        Z_\Lambda^{-1}\ee^{-\beta H_\Lambda(s)}
        \)\\[2pt]
        \(\displaystyle
        H_\Lambda=-J\sum_{\langle x y\rangle}s_xs_y-h\sum_xs_x
        \)};
    \end{scope}

    \begin{scope}[shift={(9.5,0)}]
      \node[font=\small] at (0,3.2) {diagonal trace notation};
      \node[draw=black, rounded corners, align=center, inner sep=7pt] at (0,1.55)
        {\(\displaystyle
        \Hilb_\Lambda=\bigotimes_{x\in\Lambda}V_x
        \)\\[2pt]
        \(\displaystyle
        Z_\Lambda=\operatorname{Tr}_{\Hilb_\Lambda}
        \ee^{-\beta\hat H_\Lambda}
        \)};
      \node[align=center, blue!70!black] at (0,-0.65)
        {all \(\hat S_x\) commute};
    \end{scope}

    \draw[-{Latex[length=2mm]}, thick] (3.55,2.62)--(4.35,2.62);
    \draw[-{Latex[length=2mm]}, thick] (6.9,2.62)--(7.65,2.62);
  \end{tikzpicture}
  }
  \caption{The statistical Ising model is a probability measure on lattice spin
  configurations.  Diagonal operators provide an exact finite-volume trace
  notation for the same measure.}
  \label{fig:volume-ii-ising-ensemble}
\end{figure}

For a finite set of energy eigenstates \(n\) with energies \(E_n\), the
canonical weights
\[
  p_n=\frac{1}{Z}\ee^{-\beta E_n},
  \qquad
  Z=\sum_n\ee^{-\beta E_n},
\]
are obtained by extremizing the Shannon entropy
\[
  S[p]=-\sum_n p_n\log p_n
\]
subject to \(\sum_np_n=1\) and \(\sum_np_nE_n=U\).  Indeed, the stationary
point of
\[
  S[p]+\alpha\Bigl(\sum_np_n-1\Bigr)
  +\beta\Bigl(U-\sum_np_nE_n\Bigr)
\]
has \(\log p_n=\alpha-1-\beta E_n\), hence the displayed Boltzmann form.

The entropy also has a simple monotonicity statement for symmetric probability
evolution.  If a finite distribution \(p_n(t)\) obeys
\[
  \dot p_n=\sum_m v_{nm}(p_m-p_n),
  \qquad
  v_{nm}=v_{mn}\geq0 ,
\]
then \(\sum_n\dot p_n=0\) and
\[
  \frac{\dd S}{\dd t}
  =
  \frac12\sum_{m,n}v_{mn}(p_m-p_n)(\log p_m-\log p_n)\geq0 .
\]
This is the finite-state entropy-increase mechanism underlying thermal
equilibration in the idealized symmetric-rate case.

\section{Spin Correlators and Phases}

\begin{theorem}[Ferromagnetic thermodynamic limits and phase selection]
\label{thm:ising-ferromagnetic-thermodynamic-limits}
Consider the finite-volume Ising model on boxes
\(\Lambda\nearrow\mathbb Z^D\) with finite-range couplings
\(J_{xy}=J_{yx}\geq0\), inverse temperature \(\beta\geq0\), and external
fields \(h_x\).  Let \(s_A:=\prod_{x\in A}s_x\) for finite
\(A\subset\mathbb Z^D\).  In finite volume the ferromagnetic measure satisfies
the following correlation inequalities.
\begin{enumerate}[label=(\roman*)]
  \item FKG positive association: if \(F\) and \(G\) are increasing functions
  of the spin configuration, then
  \[
    \vev{FG}_{\Lambda,\beta,h}
    \geq
    \vev{F}_{\Lambda,\beta,h}\vev{G}_{\Lambda,\beta,h}.
  \]
  \item The GKS inequalities, for nonnegative fields: for finite
  \(A,B\),
  \[
    \vev{s_A}_{\Lambda,\beta,h}\geq0,
    \qquad
    \vev{s_A s_B}_{\Lambda,\beta,h}
    -
    \vev{s_A}_{\Lambda,\beta,h}\vev{s_B}_{\Lambda,\beta,h}
    \geq0 .
  \]
  Equivalently, spin correlations are monotone nondecreasing under increasing
  ferromagnetic couplings and nonnegative external fields.
  \item The GHS inequality, for nonnegative fields:
  \[
    \frac{\partial^3\log Z_\Lambda}
      {\partial h_x\,\partial h_y\,\partial h_z}
    \leq0 .
  \]
  In particular, the finite-volume magnetization is concave as a function of
  a uniform positive field.
\end{enumerate}
Consequently the plus and minus boundary-condition Gibbs states
\[
  \vev{F}^{\pm}_{\beta,h}
  :=
  \lim_{\Lambda\nearrow\mathbb Z^D}
  \vev{F}_{\Lambda,\beta,h}^{\pm}
\]
exist for every local observable \(F\) whenever the boxes increase through a
cofinal rectangular sequence.  At \(h=0\), the one-sided states
\[
  \vev{F}^{\pm}_{\beta,0}
  =
  \lim_{h\downarrow0}\vev{F}^{\pm}_{\beta,\pm h}
\]
exist by monotonicity.  If \(\vev{\cdot}^{+}_{\beta,0}
=\vev{\cdot}^{-}_{\beta,0}\), the thermodynamic limit is independent of the
usual plus/minus/free boundary choices.  If the two differ, the plus and minus
limits are distinct translation-invariant Gibbs states.  They are pure phases
whenever they satisfy the clustering condition
\[
  \lim_{|a|\to\infty}
  \left(
    \vev{F\,\tau_aG}^{\pm}_{\beta,0}
    -
    \vev{F}^{\pm}_{\beta,0}\vev{G}^{\pm}_{\beta,0}
  \right)
  =0
\]
for all local \(F,G\), where \(\tau_a\) translates \(G\).  Clustering is the
condition that makes the limiting Gibbs state extremal, hence pure.
\end{theorem}

\begin{proof}
For the FKG assertion, order configurations by \(s\leq t\) when
\(s_x\leq t_x\) for all \(x\).  The one-edge factor
\(\exp(\beta J_{xy}s_xs_y)\) is log-supermodular for \(J_{xy}\geq0\): on the
four possible values at an edge, the only nontrivial check is
\[
  \ee^{\beta J_{xy}}\ee^{\beta J_{xy}}
  \geq
  \ee^{-\beta J_{xy}}\ee^{-\beta J_{xy}} .
\]
One-site field factors are multiplicative and hence preserve
log-supermodularity.  The full finite-volume Boltzmann weight is a product of
these factors, so it is log-supermodular, and the finite distributive-lattice
FKG theorem gives positive association.

The GKS and GHS inequalities are finite-volume correlation inequalities for
ferromagnetic Ising measures.  GKS gives nonnegativity of the derivatives of
spin correlations with respect to ferromagnetic couplings and nonnegative
fields; GHS gives the displayed concavity of the magnetization in a uniform
positive field.  These inequalities imply stochastic monotonicity in boundary
condition, coupling, and field.  Therefore the finite-volume expectations of
local observables under plus and minus boundary conditions form bounded
monotone nets, hence converge.  Any local observable is a finite linear
combination of spin monomials, so convergence of the monomial expectations
gives convergence of all local observables.

The one-sided \(h\downarrow0\) limits are again bounded monotone limits.
Equality of the plus and minus limits forces every Gibbs limit squeezed
between them by stochastic domination to agree with the common state, giving
boundary-condition independence in the usual ferromagnetic class.  If the
limits differ, they are distinct Gibbs states selected by plus and minus
boundaries, or equivalently by the sign of a symmetry-breaking field before
the field is taken to zero.  Finally, clustering implies extremality: if a
translation-invariant Gibbs state decomposed nontrivially into a convex
combination of two distinct Gibbs states, correlations of distant translates
would retain the variance of the conditional expectation onto the tail
sigma-algebra and clustering would fail.  Thus a clustering plus or minus
state is pure.
\end{proof}

The basic observables are spin correlation functions.  At \(h=0\), for
translation-invariant infinite-volume limits supplied by
Theorem~\ref{thm:ising-ferromagnetic-thermodynamic-limits}, write
\[
  C_\beta(x_1,\ldots,x_n)
  =
  \vev{s_{x_1}\cdots s_{x_n}}_\beta .
\]
The two-point function at large separation is often summarized by
\[
  f_\beta(r)
  \simeq
  \vev{s_0s_x}_\beta,
  \qquad r=|x|.
\]
For \(D=2,3\) and ferromagnetic nearest-neighbor coupling, there is a critical
temperature \(T_c\).  In the ordered region \(T<T_c\), when the plus and minus
limits are clustering and therefore pure in the sense of
Theorem~\ref{thm:ising-ferromagnetic-thermodynamic-limits}, boundary
conditions or the one-sided field limits \(h\to0^\pm\) select states with
nonzero magnetization
\[
  m_\pm(T)
  =
  \lim_{h\to0^\pm}\lim_{\Lambda\nearrow\mathbb Z^D}
  \vev{s_0}_{\Lambda,\beta,h},
  \qquad m_+(T)=-m_-(T)\neq0 .
\]
The full two-point function tends to \(m_\pm(T)^2\) at large distance in such a
phase.  The connected correlator
\[
  G_\beta(x)
  =
  \vev{s_0s_x}_\beta-\vev{s_0}_\beta\vev{s_x}_\beta
\]
decays at large \(|x|\) away from criticality.  In the disordered region
\(T>T_c\),
\[
  G_\beta(x)\sim A(T)\,\ee^{-|x|/\xi(T)}
\]
up to powers of \(|x|\), where \(\xi(T)\) is the correlation length.  One
precise definition, when the limit exists along a coordinate direction, is
\[
  \xi(T)^{-1}
  =
  -\lim_{r\to\infty}\frac1r\log |G_\beta(r e_1)|.
\]
At \(T=T_c\), the correlation length diverges and the spin two-point function
has power-law asymptotics
\[
  \vev{s_0s_x}_{\beta_c}
  \sim
  \frac{A_\sigma}{|x|^{2\Delta_\sigma}} .
\]
The exponent \(\Delta_\sigma\) is the scaling dimension of the continuum spin
field.

Figure~\ref{fig:volume-ii-ising-correlator-regimes} displays the three
large-distance regimes encoded by the two-point function: exponential decay
away from criticality, power-law behavior at criticality, and the associated
divergence of the correlation length.

\begin{figure}[H]
  \centering
  \resizebox{0.92\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \draw[-{Latex[length=2mm]}, thick] (0,0)--(7.2,0) node[below] {\(r\)};
    \draw[-{Latex[length=2mm]}, thick] (0,0)--(0,3.1) node[left] {\(f_\beta(r)\)};
    \draw[green!50!black, very thick]
      (0,2.45)..controls (1.4,2.15) and (2.7,1.95)..(4.1,1.92)
      ..controls (5.0,1.90) and (6.0,1.95)..(6.7,1.96);
    \draw[blue!75!black, very thick, dash pattern=on 5pt off 2.5pt]
      (0,2.45)..controls (1.2,1.7) and (2.4,1.10)..(3.8,0.72)
      ..controls (5.0,0.43) and (6.0,0.28)..(6.7,0.20);
    \draw[purple!75!black, very thick, densely dotted]
      (0,2.45)..controls (0.5,1.45) and (0.95,0.55)..(1.75,0.22)
      ..controls (2.5,0.06) and (3.4,0.03)..(4.8,0.02);
    \node[green!50!black, fill=white, inner sep=1pt] at (5.2,2.25) {\(T<T_c\)};
    \node[blue!75!black, fill=white, inner sep=1pt] at (5.1,0.75) {\(T=T_c\)};
    \node[purple!75!black, fill=white, inner sep=1pt] at (2.3,0.62) {\(T>T_c\)};
    \node[green!50!black, align=center, fill=white, inner sep=1pt] at (2.2,2.35)
      {ordered\\plateau};
    \node[blue!75!black, align=center, fill=white, inner sep=1pt] at (5.7,1.28)
      {\(f_\beta(r)\sim r^{-2\Delta_\sigma}\)};
    \node[purple!75!black, align=center, fill=white, inner sep=1pt] at (4.55,0.45)
      {\(\exp[-r/\xi(T)]\)};
  \end{tikzpicture}
  }
  \caption{Long-distance spin correlations distinguish the ordered,
  critical, and disordered regimes.  The ordered, critical, and disordered
  curves are drawn solid, dashed, and dotted respectively.  In the ordered
  phase the full two-point function approaches the square of the magnetization;
  the connected correlator still decays away from the critical point.}
  \label{fig:volume-ii-ising-correlator-regimes}
\end{figure}

The values relevant for the first Ising universality classes are
\[
\begin{array}{c|cc}
  D & \Delta_\sigma & \nu \\ \hline
  2 & \frac18 & 1\\[2pt]
  3 & 0.518\ldots & 0.62998\ldots
\end{array}
\]
where \(\nu\) is defined by the divergence
\[
  \xi(T)\sim \xi_0\,|T-T_c|^{-\nu}
  \qquad (T\to T_c).
\]

\section{The Scaling Limit}

The continuum Ising field theory is defined, when the limit exists, from
large-distance lattice correlators near criticality.  Let \(a=L^{-1}\) be the
lattice spacing in continuum units.  A fixed point \(x\in\mathbb R^D\)
corresponds to a nearest lattice point \([x/a]=[Lx]\).  In this subsection set
the magnetic field to zero, so the \(\mathbb Z_2\) symmetry is preserved and
the thermal coordinate is the relevant coordinate being varied.

\begin{table}[H]
\centering
\footnotesize
\begin{tabular}{@{}>{\raggedright\arraybackslash}p{0.10\textwidth}
  >{\raggedright\arraybackslash}p{0.30\textwidth}
  >{\raggedright\arraybackslash}p{0.25\textwidth}
  >{\raggedright\arraybackslash}p{0.25\textwidth}@{}}
\hline
Dimension & Ising scaling-limit status & Scalar \(\phi^4\) comparison &
Use in this chapter \\
\hline
\(D=2\) &
Critical planar Ising spin correlations and fermionic observables have
rigorously constructed conformally covariant scaling limits in standard
planar domains. &
\(P(\phi)_2\), including \(\phi^4_2\), is a constructive scalar QFT class
after Wick ordering and the required local renormalizations. &
Two-dimensional Ising may be cited as a theorem-level example of a
nontrivial lattice-to-continuum critical limit, while keeping distinct the
exact Ising lattice construction and the \(P(\phi)_2\) constructive family. \\
\(D=3\) &
The nontrivial nearest-neighbor Ising critical scaling limit and its
identification with the three-dimensional Ising CFT are not presently
available as a complete theorem with OS/Wightman reconstruction data and
universal correlators. &
Massive \(\phi^4_3\) with positive quartic interaction is a constructed
Glimm--Jaffe superrenormalizable scalar QFT: its Euclidean Schwinger
functions satisfy the OS hypotheses, hence give a Wightman QFT by
reconstruction. &
Statements about the three-dimensional Ising continuum theory are therefore
conditional on scaling-limit and universality hypotheses, even when supported
by RG, bootstrap, Monte Carlo, and perturbative evidence.  The proved massive
\(\phi^4_3\) Wightman construction is not, by itself, a construction of the
critical Wilson--Fisher/Ising fixed point. \\
\(D=4\) &
For the standard reflection-positive critical nearest-neighbor Ising scaling
problem covered by the marginal triviality theorem, scaling limits are
Gaussian. &
The same Aizenman--Duminil-Copin theorem covers the corresponding lattice
\(\phi^4_4\) scaling problem with positive quartic interaction. &
There is no interacting four-dimensional Ising/\(\phi^4_4\) continuum fixed
point from this standard lattice construction.  This theorem is distinct from
the separate logical question of nonstandard UV completions with identical
formal perturbation theory. \\
\(D>4\) &
Standard critical Ising scaling limits in the theorem classes governed by
mean-field/triviality results are Gaussian. &
Scalar \(\phi^4_D\) with positive quartic coupling is trivial in the
corresponding reflection-positive scaling-limit regimes. &
The Ising fixed point used below is not a nontrivial \(D>4\) continuum QFT;
the Wilson--Fisher deformation is meaningful below four dimensions. \\
\hline
\end{tabular}
\caption{Dimension-dependent theorem status of Ising and scalar
\(\phi^4\) continuum-limit claims.  The table separates actual
construction/triviality theorems from the conditional three-dimensional
scaling-limit assertions used in the RG discussion.}
\label{tab:ising-phi4-continuum-status}
\end{table}

\emph{References for Table~\ref{tab:ising-phi4-continuum-status}.}
For the planar Ising entries, see Smirnov's discrete-holomorphic fermion
construction for the critical random-cluster Ising model and the
Chelkak--Hongler--Izyurov proof of conformal invariance of critical spin
correlations.  For the constructive scalar entries, see the Nelson,
Simon, and Glimm--Jaffe constructions of \(P(\phi)_2\) and \(\phi^4_3\).
For the four-dimensional marginal theorem, see Aizenman--Duminil-Copin,
``Marginal triviality of the scaling limits of critical 4D Ising and
\(\phi^4_4\) models,'' \emph{Ann. of Math.} \textbf{194} (2021)
163--235, together with its corrigendum.  Earlier Aizenman and
Fr\"ohlich triviality theorems cover complementary above-four-dimensional
scalar/Ising regimes.

The \(D\geq4\) Gaussian entries in
Table~\ref{tab:ising-phi4-continuum-status} concern separated critical
scaling limits and the endpoint of the tuned RG trajectory.  They do not mean
that the quartic coordinate may be deleted from every finite-regulator
observable before the scaling limit is taken.  In the terminology of
Section~\ref{sec:dangerously-irrelevant-coordinates}, the quartic coordinate
above four dimensions is dangerously irrelevant for the ordered equation of
state: it flows to zero at the Gaussian fixed point, but it stabilizes the
ordered saddle and appears singularly in the magnetization and free-energy
amplitudes.  At \(D=4\) the same coordinate is marginally irrelevant and
produces the leading logarithmic corrections derived in
\eqref{eq:dangerous-four-dimensional-xi-log} and
\eqref{eq:dangerous-four-dimensional-magnetization-log}.  This separates the
Gaussian character of the critical separated-point continuum limit from the
singular quartic dependence of ordered thermodynamic observables.

Let \(\xi_{\rm lat}(T)\) be the correlation length measured in lattice units,
using connected correlators in a chosen pure phase when \(T<T_c\).  Choose a
sequence of temperatures \(T(L)\to T_c\) such that the signed thermal scale
\[
  \mathfrak m
  =
  \lim_{L\to\infty}
  \operatorname{sgn}(T(L)-T_c)\,
  \frac{L}{\xi_{\rm lat}(T(L))}
\]
exists.  The physical inverse correlation length is \(|\mathfrak m|\).  The
sign records which side of the critical point is approached.  More general
scaling limits with a magnetic deformation require a second scaling coordinate
to be held fixed.  If \(u_h\) is the coefficient of the spin field, this
coordinate has the form \(L^{D-\Delta_\sigma}u_h\).  It is set to zero here.
The quantity \(|\mathfrak m|\) is the continuum mass gap extracted from the
lattice transfer-matrix or correlation-length scale.  It is a spectral pole
mass only under the additional condition that the reconstructed two-point
spectral measure in the relevant channel has an isolated lowest one-particle
atom, as in Definition~\ref{def:mass-data-and-mass-coordinates}; otherwise it
is the lower edge of the spectrum in that channel.

\begin{hypothesis}[Separated spin scaling limit]
\label{hyp:ising-separated-spin-scaling-limit}
For the temperature trajectory above, assume that there is a spin-field
renormalization factor \(Z_\sigma(L,T)>0\) such that, for every \(n\) and every
configuration of pairwise distinct continuum points \(x_1,\ldots,x_n\), the
limit
\[
  \lim_{\substack{T\to T_c,\ L\to\infty\\
      \operatorname{sgn}(T-T_c)L/\xi_{\rm lat}(T)\to\mathfrak m}}
  Z_\sigma(L,T)^{-n/2}
  \vev{s_{[Lx_1]}\cdots s_{[Lx_n]}}_{\beta(T)}
  =
  \vev{\sigma(x_1)\cdots\sigma(x_n)}_{\operatorname{Ising},\mathfrak m}
\]
exists.  The convergence is meant after the thermodynamic limit specified for
the lattice model, and with the boundary/pure-phase convention fixed when
\(\mathfrak m<0\).
\end{hypothesis}

When Hypothesis~\ref{hyp:ising-separated-spin-scaling-limit} holds, its right
side defines the separated-point spin Schwinger functions of the continuum
Ising scaling theory at thermal scale \(\mathfrak m\).  At
\(\mathfrak m=0\), equivalently at \(\xi_{\rm lat}(T)/L\to\infty\), the
limiting fixed point is called the critical Ising fixed point.  If this fixed
point admits a local stress tensor that is conserved, symmetric, and
improvable to a traceless tensor whose charges generate the conformal
transformations of the separated-point Schwinger functions, then the fixed
point is the Ising conformal field theory.  At criticality
\(Z_\sigma(L,T_c)\) may be chosen proportional to \(L^{-2\Delta_\sigma}\), so
the rescaled spin insertions have finite correlators.
The limiting objects in this display are Euclidean separated-point Schwinger
functions.  Reflection positivity of these limiting distributions is not a
consequence of the scaling formula alone.  It is a property that must either be
proved from a reflection-positive regulated lattice model and a sufficiently
strong mode of convergence, or imposed as a separate reconstruction
hypothesis.  When this positivity and the remaining reconstruction hypotheses
hold, the Schwinger functions determine a Lorentzian QFT Hilbert space by the
Osterwalder--Schrader reconstruction of
Chapter~\ref{ch:os-reconstruction}.  That reconstructed Hilbert space is not
the finite-dimensional spin-trace space \(\Hilb_\Lambda\) used to define the
classical thermal ensemble.

Figure~\ref{fig:volume-ii-ising-scaling-limit} records the scaling-limit map:
renormalized lattice insertions converge to separated-point Schwinger
functions, and OS reconstruction is an additional step after reflection
positivity and the remaining hypotheses are established.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small] at (1.7,3.05) {lattice insertions};
      \draw[step=0.55, gray!45] (0,0) grid (3.3,2.2);
      \foreach \x/\y in {0.55/1.65,1.65/1.1,2.75/0.55} {
        \fill[blue!70!black] (\x,\y) circle (2.4pt);
      }
      \node[align=center] at (1.65,-0.45)
        {\([Lx_i]\) nearest lattice points};
    \end{scope}

    \draw[-{Latex[length=2mm]}, thick] (3.9,1.1)--(5.1,1.1)
      node[midway, above, align=center] {\(L\to\infty\)\\\(T\to T_c\)};

    \begin{scope}[shift={(5.8,0)}]
      \node[font=\small] at (1.65,3.05) {continuum fields};
      \draw[-{Latex[length=2mm]}, gray!70] (0,0)--(3.35,0) node[below] {\(x^1\)};
      \draw[-{Latex[length=2mm]}, gray!70] (0,0)--(0,2.35) node[left] {\(x^2\)};
      \foreach \x/\y in {0.55/1.65,1.65/1.1,2.75/0.55} {
        \fill[purple!75!black] (\x,\y) circle (2.4pt);
        \node[purple!75!black, above right=-1pt] at (\x,\y) {\(\sigma(x_i)\)};
      }
      \node[align=center, blue!70!black] at (1.65,-0.45)
        {\(L/\xi_{\rm lat}(T)\to |\mathfrak m|\)};
    \end{scope}

    \node[draw=black, rounded corners, align=center, inner sep=7pt]
      at (5.0,-1.45)
      {\(\displaystyle
        Z_\sigma^{-n/2}\,
        \vev{s_{[Lx_1]}\cdots s_{[Lx_n]}}
        \longrightarrow
        \vev{\sigma(x_1)\cdots\sigma(x_n)}_{\operatorname{Ising},\mathfrak m}
      \)};
  \end{tikzpicture}
  }
  \caption{The Ising field theory is the continuum scaling limit of
  long-distance spin correlators.  The absolute value of \(\mathfrak m\) is
  the inverse correlation length in the limiting continuum units, while its
  sign records the high- or low-temperature side.}
  \label{fig:volume-ii-ising-scaling-limit}
\end{figure}

\section{Reflection Positivity Along the Scaling Limit}

Choose one lattice coordinate and call it Euclidean time.  Let
\(\theta\) be reflection across the hyperplane midway between the time slices
\(n^0=0\) and \(n^0=1\):
\[
  \theta(n^0,n^1,\ldots,n^{D-1})
  =
  (1-n^0,n^1,\ldots,n^{D-1}) .
\]
For a finite reflection-invariant lattice region \(\Lambda\), write
\(\Lambda_+=\{n\in\Lambda:n^0\geq1\}\), and let
\(\mathcal A_+\) be the algebra of complex functions of the lattice variables
with support in \(\Lambda_+\).  If \(F\in\mathcal A_+\), define the
anti-linear reflection
\[
  (\Theta F)(q)=\overline{F(q\circ\theta)} ,
\]
where \(q\) denotes the complete lattice configuration.

\begin{definition}[Finite-lattice reflection positivity]
A finite lattice probability measure \(\mu_\Lambda\) is reflection positive
with respect to the time reflection \(\theta\) if
\begin{equation}
  \int (\Theta F)(q)\,F(q)\,\dd\mu_\Lambda(q)\geq0
  \label{eq:finite-lattice-reflection-positivity}
\end{equation}
for every \(F\in\mathcal A_+\).  Equivalently, the sesquilinear form
\((F,G)\mapsto\int(\Theta F)G\,\dd\mu_\Lambda\) is positive semidefinite on
\(\mathcal A_+\).
\end{definition}

This is the finite-cutoff precursor of the Osterwalder--Schrader positivity
condition.  It is stronger than pointwise positivity of the Boltzmann weight:
it constrains how the action couples fields on opposite sides of every
reflection plane used in reconstruction.

\begin{proposition}[Ferromagnetic nearest-neighbor models are reflection positive]
\label{prop:ferromagnetic-nearest-neighbor-reflection-positive}
Consider a reflection-invariant finite box with free boundary conditions, or a
periodic torus with the reflection plane chosen so that the action decomposes
into reflected halves and bonds crossing the plane.  The source-free
nearest-neighbor Ising measure with coupling \(J\geq0\) is reflection
positive.  The real scalar lattice measure
\begin{equation}
  \dd\mu_\Lambda(s)
  =
  Z_\Lambda^{-1}
  \prod_{x\in\Lambda}\rho(\dd s_x)\,
  \exp\!\left[
    \sum_{\langle x y\rangle}K\,s_xs_y
  \right],
  \qquad K\geq0,
  \label{eq:scalar-ferromagnetic-rp-measure}
\end{equation}
is also reflection positive whenever \(\rho\) is a positive one-site measure
for which the finite-volume partition function exists.  Thus the
nearest-neighbor generalized Ising model with
\(\rho(\dd s)=\exp[-P(s)]\dd s\), \(P\) bounded below with sufficient growth,
and \(K=\beta\geq0\), preserves reflection positivity.  Convexity of \(P\) is
not required; a double-well one-site potential does not by itself violate
reflection positivity.
\end{proposition}

\begin{proof}
The Ising model is the special case
\(\rho=\frac12(\delta_{-1}+\delta_{+1})\) and \(K=\beta J\).  It is therefore
enough to prove the scalar statement.  Decompose the action into terms
strictly on the positive side, their reflected copies, and the bonds crossing
the reflection plane:
\[
  \exp[-S_\Lambda(s)]
  =
  \exp[-S_+(s_+)]\,
  \exp[-S_+(\theta s_-)]\,
  \prod_{y\in\Sigma_+}\exp\!\left[K_y s_{\theta y}s_y\right],
\]
where \(y\) runs over the positive-side endpoints of the crossing bonds and
\(K_y\geq0\).  Expanding every crossing factor gives
\[
  \prod_{y\in\Sigma_+}\exp[K_y s_{\theta y}s_y]
  =
  \sum_{m:\Sigma_+\to\mathbb Z_{\geq0}}
  \prod_{y\in\Sigma_+}\frac{K_y^{m_y}}{m_y!}\,
  \prod_{y\in\Sigma_+}s_{\theta y}^{m_y}
  \prod_{y\in\Sigma_+}s_y^{m_y}.
\]
Substitution into the quadratic form
\(\int(\Theta F)F\,\dd\mu_\Lambda\) factorizes the positive and negative
half-lattice integrals.  For each multi-index \(m\) the contribution is
\[
  \frac{1}{Z_\Lambda}
  \prod_{y\in\Sigma_+}\frac{K_y^{m_y}}{m_y!}
  \left|
    \int
    F(s_+)\,
    \prod_{y\in\Sigma_+}s_y^{m_y}\,
    \exp[-S_+(s_+)]
    \prod_{x\in\Lambda_+}\rho(\dd s_x)
  \right|^2 .
\]
The coefficients are nonnegative because \(K_y\geq0\).  The sum of these
nonnegative terms is the desired quadratic form, proving
\eqref{eq:finite-lattice-reflection-positivity}.
\end{proof}

The proposition identifies a sufficient condition, not a definition of the
universality class.  Adding next-nearest-neighbor, anisotropic, or
higher-derivative lattice couplings preserves reflection positivity only when
the interaction crossing each reflection plane is a positive kernel in the
above sense.  Negative cross-reflection couplings, or derivative
discretizations whose cross-plane kernel has negative directions, generally
destroy the factorization into sums of squares.  Such models may still have
the same formal RG relevant coordinates, but they do not automatically supply
Osterwalder--Schrader positivity for a Lorentzian reconstruction.

\begin{proposition}[Closedness of reflection positivity under scaling limits]
\label{prop:reflection-positivity-closed-under-scaling-limits}
Let \(a\downarrow0\) label reflection-positive lattice measures
\(\mu_a\).  Let \(A_a^I(f)\) be renormalized smeared lattice observables whose
linear normalizations are real and whose test functions are supported in the
positive Euclidean half-space.  Suppose that for every finite collection of
polynomial functionals \(F_\alpha(A_a)\) supported in that half-space and
every set of complex coefficients \(c_\alpha\), the reflected quadratic forms
\[
  Q_a(c)
  =
  \sum_{\alpha,\beta}\overline{c_\alpha}c_\beta\,
  \left\langle
    \Theta F_\alpha(A_a)\,F_\beta(A_a)
  \right\rangle_{\mu_a}
\]
converge to the corresponding continuum quadratic form \(Q_\ast(c)\).  Then
\(Q_\ast(c)\geq0\) for all such \(c_\alpha\); the limiting Schwinger
distributions are reflection positive on this test-function domain.
\end{proposition}

\begin{proof}
For every \(a\), finite-lattice reflection positivity gives
\(Q_a(c)\geq0\).  The assumed convergence of the quadratic forms gives
\[
  Q_\ast(c)=\lim_{a\downarrow0}Q_a(c),
\]
and the closedness of the cone \([0,\infty)\subset\mathbb R\) gives
\(Q_\ast(c)\geq0\).
\end{proof}

The hypothesis in this proposition concerns reflected quadratic forms and
separated-point correlators together.  Source-local
subtractions used to define a distributional generating functional must either
arise from genuine lattice observables whose reflected quadratic forms
converge, or be separately checked to preserve the Osterwalder--Schrader
quadratic inequalities.  Coordinate universality and critical exponent
matching do not imply this positivity statement.

Microscopic lattice observables become continuum operators through their
leading terms in the scaling-field expansion.  Let
\[
  e_x
  =
  -\sum_{\mu=1}^D s_xs_{x+\hat\mu}
\]
be a local energy-density observable on the hypercubic lattice.  At separated
continuum points and at criticality,
\[
  s_{[x/a]}
  =
  C_\sigma\,a^{\Delta_\sigma}\sigma(x)
  +
  \sum_j C_j\,a^{\Delta_j}O_j(x),
  \qquad
  \Delta_j>\Delta_\sigma ,
\]
and
\[
  e_{[x/a]}-\vev{e}_{c}
  =
  C_\varepsilon\,a^{\Delta_\varepsilon}\varepsilon(x)
  +
  \sum_k C'_k\,a^{\Delta'_k}O'_k(x),
  \qquad
  \Delta'_k>\Delta_\varepsilon .
\]
The constants \(C_\sigma\) and \(C_\varepsilon\) depend on the microscopic
normalization of the lattice observables.  The dimensions and operator-product
coefficients of the scaling fields are fixed-point data.

In scalar continuum coordinates, the leading \(\mathbb Z_2\)-odd scalar field
\(\phi\) represents the Ising spin field \(\sigma\), while the leading
\(\mathbb Z_2\)-even mass operator, denoted \(\phi^2\) after identity mixing is
subtracted, represents the energy field \(\varepsilon\):
\[
  \phi \longrightarrow \sigma,
  \qquad
  [\phi^2]\longrightarrow\varepsilon .
\]
If \(u_t\) denotes the thermal scaling coordinate, then
\[
  \Delta S_t=\int\dd^D x\,u_t\,\varepsilon(x),
  \qquad
  [u_t]=D-\Delta_\varepsilon .
\]
Since \(\xi^{-1}\propto |u_t|^{1/(D-\Delta_\varepsilon)}\), comparison with
\(\xi\propto |T-T_c|^{-\nu}\) gives
\[
  \Delta_\varepsilon
  =
  D-\frac1\nu .
\]

\section{Observable Limits as Source-Dependent Distributions}

The separated-point scaling formula is the restriction of a stronger
distributional statement.  Let \(a\) be the lattice spacing and let
\(\mathcal D(\mathbb R^D)\) be the space of smooth compactly supported test
functions.  For \(f,g\in\mathcal D(\mathbb R^D)\), define the renormalized
smeared lattice observables
\[
  \Sigma_a(f)
  =
  a^D\sum_{n\in\mathbb Z^D}
  f(an)\,C_\sigma^{-1}a^{-\Delta_\sigma}s_n,
\]
and
\[
  E_a(g)
  =
  a^D\sum_{n\in\mathbb Z^D}
  g(an)\,C_\varepsilon^{-1}a^{-\Delta_\varepsilon}
  \bigl(e_n-\vev{e}_c\bigr),
\]
where the sums are restricted to the finite lattice region before the
thermodynamic limit is taken.  The constants \(C_\sigma\) and
\(C_\varepsilon\) are the same nonuniversal normalization constants appearing
in the leading scaling-field expansion.  The point fields \(\sigma(x)\) and
\(\varepsilon(x)\) are the distributional kernels whose smeared limits are
\(\Sigma(f)=\int\dd^Dx\,f(x)\sigma(x)\) and
\(E(g)=\int\dd^Dx\,g(x)\varepsilon(x)\).

\begin{definition}[Source-dependent scaling limit]
Fix the thermal scaling trajectory used above.  A source-dependent scaling
limit for a chosen collection of microscopic local observables \(A_a^I\)
consists of a continuum source coordinate \(\eta^A(x)\), a lattice source
coordinate \(\eta_a^I(n)\), and local source maps
\[
  \eta_a=\mathcal Z_a(\eta),
\]
such that, after subtraction of source-local vacuum terms,
\begin{equation}
  W_a[\mathcal Z_a(\eta)]-W_a[0]-P_a(\eta)
  \longrightarrow
  W_\ast[\eta]
  \label{eq:ising-source-dependent-scaling-limit}
\end{equation}
as a functional of test sources near \(\eta=0\).  Here
\[
  \exp W_a[\eta_a]
  =
  \left\langle
    \exp\left[
      -a^D\sum_n\eta_a^I(n)A_a^I(n)
    \right]
  \right\rangle_a ,
\]
and \(P_a(\eta)\) is a local polynomial in the continuum sources and their
derivatives.  The linear part of \(\mathcal Z_a\) fixes field normalization
and operator mixing.  Its higher local Taylor coefficients fix the contact
extension of products of continuum fields.
\end{definition}

The term \(P_a(\eta)\) removes vacuum terms such as powers of the volume and
local source self-contractions.  It has no effect on separated normalized
correlators, but it is part of the definition of the full distributional
generating functional.

\begin{proposition}[Separated universality]
Assume the source-dependent scaling limit
\eqref{eq:ising-source-dependent-scaling-limit} exists for two microscopic
realizations \(X\) and \(Y\) and that their continuum source coordinates have
been identified by the same fixed-point operator normalization.  Then their
noncoincident continuum correlators of the corresponding scaling fields agree.
Their contact-term extensions may differ by local distributions supported on
collision diagonals.
\end{proposition}

\begin{proof}
Functional derivatives of
\eqref{eq:ising-source-dependent-scaling-limit} at \(\eta=0\) give the
scaling limits of smeared inserted correlators.  Restricting the resulting
distributions to the configuration space of distinct insertion points removes
all terms supported on collision diagonals.  By the source-coordinate
partition formula of
Section~\ref{sec:multiple-insertions-contact-coordinates}, only the linear
source jet contributes on this configuration space.  Once the linear jet has
been fixed by the common operator normalization, the two realizations have the
same separated-point correlators.  Higher source jets and the local
polynomials \(P_a\) change only the extension across collision diagonals.
\end{proof}

Thus microscopic observables enter the continuum theory through renormalized
source derivatives whose limits are operator-valued distributions.  The
universal data are the separated-point kernels obtained after the relevant
scaling coordinates and linear operator normalizations have been fixed.  These
kernels are the source-coordinate input for the fixed-point operator data
constructed in Chapter~\ref{ch:volume-iii-fixed-points-conformal-data}.

\section{A Lattice Scalar Path Integral}

The Ising variables may be enlarged from two values to a real local variable
without changing the long-distance fixed point, provided the relevant
\(\mathbb Z_2\)-even coupling is tuned.  Let \(P:\mathbb R\to\mathbb R\) be an
even single-site potential with sufficient growth at infinity.  The generalized
Ising partition function is
\[
  \widetilde Z_\Lambda
  =
  \int\prod_{x\in\Lambda}\dd s_x\,
  \exp\!\left[
    -\sum_{x\in\Lambda}P(s_x)
    +
    \beta\sum_{\langle x y\rangle}s_xs_y
  \right].
\]
For \(P\) real and bounded below with sufficient growth and for
\(\beta\geq0\), this nearest-neighbor scalar measure is in the
reflection-positive class of
Proposition~\ref{prop:ferromagnetic-nearest-neighbor-reflection-positive}.
This assertion uses the positivity of the cross-reflection nearest-neighbor
coupling.  It would not follow for an arbitrary lattice regularization with
negative next-nearest-neighbor terms or other cross-plane kernels that are not
positive in the sense used in the proof of that proposition.
Equivalently,
\[
  \widetilde Z_\Lambda
  =
  \int\prod_{x\in\Lambda}\dd s_x\,\ee^{-S_\Lambda(s)}
\]
with
\[
  S_\Lambda(s)
  =
  \frac{\beta}{2}\sum_{\langle x y\rangle}(s_x-s_y)^2
  +
  \sum_{x\in\Lambda}\bigl(P(s_x)-\beta D\,s_x^2\bigr),
\]
where each positive lattice direction is counted once.  The identity follows
from
\[
  \sum_{\langle x y\rangle}(s_x-s_y)^2
  =
  2D\sum_xs_x^2-2\sum_{\langle x y\rangle}s_xs_y
\]
on a periodic hypercubic lattice.

For an infinite hypercubic lattice with unit spacing, introduce the Brillouin
zone
\[
  B=[-\pi,\pi]^D
\]
and write
\[
  s_x
  =
  \int_B\frac{\dd^Dk}{(2\pi)^D}\,
  \widetilde\phi(k)\,\ee^{\ii k\cdot x},
  \qquad
  \widetilde\phi(k+2\pi n)=\widetilde\phi(k),
  \quad n\in\mathbb Z^D .
\]
The cutoff \(|k^\mu|\leq\pi\) is a finite ultraviolet cutoff.  The nearest
neighbor kinetic term becomes
\[
  \sum_{\langle x y\rangle}(s_x-s_y)^2
  =
  \int_B\frac{\dd^Dk}{(2\pi)^D}\,
  \widetilde\phi(-k)\widetilde\phi(k)
  \sum_{\mu=1}^D\left|\ee^{\ii k_\mu}-1\right|^2 .
\]
For lattice spacing \(a\), the dimensionful momentum is \(p=k/a\), and the
Brillouin-zone edge is therefore
\[
  \Lambda_{\rm BZ}=\frac{\pi}{a}.
\]
In the finite-cutoff Wilsonian estimates below, \(\Lambda_0=a^{-1}\) denotes
the inverse lattice spacing rather than the Brillouin-zone edge.  The two
conventions are related by the fixed conversion
\[
  \Lambda_{\rm BZ}=\pi\Lambda_0 .
\]
Replacing \(\Lambda_0\) by \(\Lambda_{\rm BZ}\) in a cutoff-suppressed estimate
only multiplies the corresponding constant by a power of \(\pi\); it does not
change the exponent or the continuum limit.
For small momentum,
\[
  \sum_{\mu=1}^D\left|\ee^{\ii k_\mu}-1\right|^2
  =
  k^2-\frac1{12}\sum_{\mu=1}^D k_\mu^4+O(k^6).
\]
Thus the long-distance action has the form
\[
  S[\phi]
  =
  \int \dd^Dx
  \left[
    \frac{\beta}{2}\sum_{\mu=1}^D(\partial_\mu\phi)^2
    +
    V_\beta(\phi)
    -
    \frac{\beta}{24}\sum_{\mu=1}^D(\partial_\mu^2\phi)^2
    +\cdots
  \right],
\]
with \(V_\beta(\phi)=P(\phi)-\beta D\phi^2\) plus local terms generated by the
field normalization convention.  The higher-derivative terms reflect the
lattice cutoff and the breaking of continuous rotations to the hypercubic
lattice symmetry.

Figure~\ref{fig:volume-ii-ising-lattice-scalar} shows the local scalar action
generated from the lattice ensemble: the single-site measure supplies
\(V_\beta(\phi)\), the nearest-neighbor term supplies the gradient term, and
higher-derivative operators record the cutoff symmetry.

\begin{figure}[H]
  \centering
  \resizebox{0.98\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small] at (0,2.85) {single-site measure};
      \draw[-{Latex[length=2mm]}, thick] (-1.4,0)--(1.45,0) node[below] {\(s\)};
      \draw[-{Latex[length=2mm]}, thick] (0,-0.25)--(0,2.15) node[left] {\(P(s)\)};
      \draw[blue!70!black, very thick]
        plot[smooth] coordinates {(-1.3,1.9) (-0.75,0.35) (0,1.15) (0.75,0.35) (1.3,1.9)};
      \node[blue!70!black] at (0,-0.55) {\(\mathbb Z_2\)-even};
    \end{scope}

    \begin{scope}[shift={(4.2,0)}]
      \node[font=\small] at (0,2.85) {cutoff momentum space};
      \draw[thick] (-1.3,-1.3) rectangle (1.3,1.3);
      \draw[-{Latex[length=2mm]}, gray!70] (-1.65,0)--(1.8,0) node[below] {\(k^1\)};
      \draw[-{Latex[length=2mm]}, gray!70] (0,-1.65)--(0,1.8) node[left] {\(k^2\)};
      \node at (-1.3,-1.55) {\(-\pi\)};
      \node at (1.3,-1.55) {\(\pi\)};
      \node[align=center, blue!70!black] at (0,-2.0)
        {Brillouin zone \(B\)};
    \end{scope}

    \begin{scope}[shift={(8.8,0)}]
      \node[font=\small] at (0,2.85) {derivative expansion};
      \node[draw=black, rounded corners, align=center, inner sep=7pt] at (0,0.85)
        {\(\displaystyle
        \frac{\beta}{2}(\partial\phi)^2+V_\beta(\phi)
        \)\\[2pt]
        \(\displaystyle
        -\,\frac{\beta}{24}\sum_\mu(\partial_\mu^2\phi)^2+\cdots
        \)};
      \node[align=center, purple!75!black] at (0,-1.05)
        {irrelevant lattice\\anisotropy};
    \end{scope}

    \draw[-{Latex[length=2mm]}, thick] (1.85,0.75)--(3.05,0.75);
    \draw[-{Latex[length=2mm]}, thick] (5.9,0.75)--(7.25,0.75);
  \end{tikzpicture}
  }
  \caption{The generalized Ising model may be rewritten as a scalar lattice
  path integral.  The single-site measure supplies the potential, the nearest
  neighbor interaction supplies the kinetic term, and the Brillouin zone gives
  a finite ultraviolet cutoff.}
  \label{fig:volume-ii-ising-lattice-scalar}
\end{figure}

\section{Universality Classes and Relevant Coordinates}

The lattice scalar action just constructed and the continuum quartic scalar
action differ in their short-distance coordinates.  The lattice action has a
finite cutoff, lattice-anisotropic derivative terms, and many allowed local
couplings such as \(\phi^6\) and higher powers or derivatives.  These
differences define distinct points in the same local theory space.  The
universality statement is that, in \(D=2,3\) and under the existence of the
scaling limit described above, these distinct ultraviolet points have
long-distance trajectories that approach the same Ising fixed point after the
relevant coordinates are tuned.

Near the fixed point, write a general local deformation as
\[
  \Delta S
  =
  \int\dd^Dx\,
  \left(
    u_h\,\sigma(x)
    +
    u_t\,\varepsilon(x)
    +
    \sum_I u_I\,O_I(x)
  \right),
  \qquad
  \Delta_I=\dim O_I .
\]
The coefficient \(u_I\) has mass dimension \(D-\Delta_I\).  At a physical mass
scale \(\mu\), the dimensionless strength of this perturbation is
\[
  \lambda_I(\mu)=u_I\,\mu^{\Delta_I-D}.
\]
To linear order at the fixed point,
\[
  \frac{\dd\lambda_I}{\dd\log\mu}
  =
  (\Delta_I-D)\lambda_I+O(\lambda^2).
\]
If \(\Delta_I>D\), the dimensionless strength tends to zero as \(\mu\to0\).
These are the irrelevant perturbations.  They account for microscopic
differences such as the choice of square or hexagonal lattice, the details of
\(P(s)\), and higher-derivative cutoff terms.

The fixed-point spectral input used in the Ising universality statement is
that, after omitting the identity, total derivatives, equation-of-motion
descendants, and metric deformations, the only relevant scalar action
deformations are the \(\mathbb Z_2\)-odd spin field \(\sigma\) and the
\(\mathbb Z_2\)-even energy field \(\varepsilon\):
\[
  \Delta_\sigma<D,
  \qquad
  \Delta_\varepsilon=D-\frac1\nu<D .
\]
Their linear RG eigenvalues are
\[
  y_h=D-\Delta_\sigma,
  \qquad
  y_t=D-\Delta_\varepsilon=\frac1\nu .
\]
The magnetic field couples to \(u_h\).  If the microscopic \(\mathbb Z_2\)
symmetry is imposed, then \(u_h=0\) and the displayed theory-space slice has
only the thermal relevant direction.  The temperature deviation determines the
thermal scaling coordinate by
\[
  u_t
  =
  c_T(T-T_c)+O((T-T_c)^2),
  \qquad
  c_T\neq0 ,
\]
after a choice of temperature coordinate near \(T_c\).  Since \(u_t\) has
dimension \(D-\Delta_\varepsilon\), it generates the correlation length
\[
  \xi(T)
  \propto
  |u_t|^{-1/(D-\Delta_\varepsilon)}
  \propto
  |T-T_c|^{-\nu}.
\]
The coefficient \(c_T\) and the amplitude of \(\xi\) are not universal; the
exponent is.

Figure~\ref{fig:volume-ii-ising-universality-rg} gives the local RG picture of
the Ising universality statement: the temperature coordinate follows the
relevant direction, and microscopic data along irrelevant directions are
forgotten in the scaling limit.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,3.6) {\(\mathbb Z_2\)-even theory space near the Ising fixed point};
    \fill[purple!80!black] (0,0) circle (3pt);

    \node[draw=black, rounded corners, align=center, fill=blue!7, inner sep=5pt]
      (phi) at (-4.2,2.2)
      {continuum scalar\\quartic actions};
    \node[draw=black, rounded corners, align=center, fill=green!7, inner sep=5pt]
      (lat) at (4.2,2.2)
      {Ising lattice\\actions};
    \node[draw=black, rounded corners, align=center, fill=gray!8, inner sep=5pt]
      (irr) at (0,2.75)
      {irrelevant local\\differences};

    \draw[-{Latex[length=2mm]}, very thick, blue!75!black]
      (phi.south east)..controls (-2.5,1.2) and (-1.3,0.55)..(-0.15,0.1);
    \draw[-{Latex[length=2mm]}, very thick, green!50!black,
      dash pattern=on 5pt off 2.5pt]
      (lat.south west)..controls (2.5,1.2) and (1.3,0.55)..(0.15,0.1);
    \foreach \x in {-1.5,-0.7,0.7,1.5} {
      \draw[-{Latex[length=2mm]}, gray!70, thick, densely dotted]
        (\x,2.25)..controls (0.6*\x,1.55) and (0.25*\x,0.8)..(0.05*\x,0.18);
    }

    \draw[-{Latex[length=2mm]}, red!75!black, very thick,
      dash pattern=on 4pt off 2pt]
      (0,0)--(0,-2.0);
    \node[red!75!black, align=center] at (1.85,-1.15)
      {temperature or mass\\deformation};
    \node[blue!70!black, align=center] at (-2.55,-1.35)
      {critical tuning\\sets this coordinate to zero};
    \draw[-{Latex[length=2mm]}, blue!70!black] (-1.8,-1.1)--(-0.2,-0.25);
    \node[purple!80!black, align=center, fill=white, inner sep=2pt,
      anchor=east]
      at (-0.25,0.72)
      {Ising CFT\\fixed point};
  \end{tikzpicture}
  }
  \caption{The \(\mathbb Z_2\)-even theory-space slice has a basin whose
  critical trajectories approach the same Ising fixed point.  Solid and dashed
  arrows distinguish continuum scalar and lattice trajectories; dotted arrows
  indicate irrelevant differences.  Relevant thermal deformations move the
  theory away from criticality and produce massive scaling limits.}
  \label{fig:volume-ii-ising-universality-rg}
\end{figure}

Thus the statistical Ising model supplies a constructive route from
microscopic lattice ensembles to continuum Euclidean QFT.  The construction is
by correlators: after tuning the relevant coordinates and rescaling fields, the
large-distance spin correlators define correlators of local fields at, or near,
the Ising fixed point.

\section{Finite-Cutoff Universality in a Local Chart}

The preceding paragraph can be made into a local finite-cutoff statement.
Let \(\Lambda_0=a^{-1}\) be the lattice cutoff and let \(\mu_R\) be a fixed
reference momentum scale with \(\mu_R<\Lambda_0\).  Choose a finite local RG
chart near the Ising fixed point and write the projected Wilsonian coordinates
at \(\mu_R\) as
\[
  z(\mu_R;\Lambda_0)=(u(\mu_R;\Lambda_0),v(\mu_R;\Lambda_0)).
\]
Here \(u\) denotes the relevant coordinates and \(v\) denotes the retained
irrelevant coordinates.  In the full scalar chart
\[
  u=(\lambda_h,\lambda_t),
\]
where \(\lambda_h\) is the dimensionless coordinate coupled to the spin field
\(\sigma\), and \(\lambda_t\) is the dimensionless coordinate coupled to the
energy field \(\varepsilon\).  In the \(\mathbb Z_2\)-even chart
\(\lambda_h\) is absent.  The target value
\[
  u_R=(\lambda_{h,R},\lambda_{t,R})
\]
specifies either the critical theory, when \(u_R=0\), or a massive scaling
limit, when one of the relevant coordinates is held at a fixed nonzero value
at the scale \(\mu_R\).

Let \(X\) label a microscopic realization in the same proposed universality
class: a nearest-neighbor Ising model, a generalized single-site potential, or
a lattice scalar action with different short-distance couplings.  Let
\(p_X\) denote the microscopic parameters available for tuning, such as
temperature and magnetic field.  The endpoint map in the finite chart is
\[
  E_X(\Lambda_0,p_X)
  =
  u_X(\mu_R;\Lambda_0,p_X).
\]
The local tuning problem is to solve
\[
  E_X(\Lambda_0,p_X)=u_R .
\]
In the \(\mathbb Z_2\)-even one-parameter case this is the precise version of
tuning the temperature: if \(p_X=T\), then the condition at criticality is
\(\lambda_t(\mu_R;\Lambda_0,T)=0\).

\begin{proposition}[Local universality from RG coordinates]
\label{prop:local-universality-rg-coordinates}
Fix a finite local RG chart near the Ising fixed point and a target relevant
coordinate \(u_R\).  Suppose that, for every sufficiently large
\(\Lambda_0\) and for each microscopic realization \(X\) under consideration,
the following conditions hold.
\begin{enumerate}[label=(\roman*)]
  \item The endpoint equation
  \(E_X(\Lambda_0,p_X)=u_R\) has a solution \(p_X(\Lambda_0,u_R)\), and the
  corresponding trajectory remains in the chosen chart between
  \(\Lambda_0\) and \(\mu_R\).
  \item The retained irrelevant block obeys the same backward contraction
  estimate used in the Wilsonian cutoff-removal theorem.
  \item The microscopic irrelevant coordinates are uniformly bounded after
  expressing the ultraviolet-regulated microscopic action in the chosen local
  coordinates.
  \item The generated integral along the tuned trajectory converges uniformly
  as \(\Lambda_0\to\infty\).
\end{enumerate}
Then there is a graph \(v=V_{\mu_R}(u_R)\), depending on the fixed point and
on the chosen RG coordinates but not on the microscopic label \(X\), such that
\[
  v_X(\mu_R;\Lambda_0,p_X(\Lambda_0,u_R))
  =
  V_{\mu_R}(u_R)
  +
  O\!\left[
    \left(\frac{\mu_R}{\Lambda_0}\right)^\kappa
  \right]
\]
for some \(\kappa>0\) determined by the irrelevant exponents and by the
convergence rate of the generated integral.
\end{proposition}

\begin{proof}
Apply Theorem~\ref{thm:conditional-finite-coordinate-wilsonian-continuum-limit}
to the RG trajectory whose relevant endpoint coordinate is fixed to be
\(u_R\).  The irrelevant part of the bare coordinate contributes at order
\((\mu_R/\Lambda_0)^\omega\), where \(\omega\) is the smallest retained
irrelevant exponent in the chart.  The generated integral has the assumed
cutoff limit and a convergence rate, say \(\omega_F\).  Taking
\(\kappa=\min(\omega,\omega_F)\) gives the displayed estimate.  The limiting
graph is independent of \(X\) because all dependence on \(X\) has been placed
either in the solved relevant-coordinate tuning or in bounded irrelevant
bare data whose direct memory is power-suppressed.
\end{proof}

This proposition is a coordinate form of universality.  It separates three
kinds of data.  First, the fixed point and its scaling operators determine
\(\Delta_\sigma\), \(\Delta_\varepsilon\), \(y_h\), \(y_t\), and the limiting
operator algebra.  Second, the microscopic model determines nonuniversal
coordinate maps, such as
\[
  \lambda_t(\Lambda_0)
  =
  A_X\,(T-T_{c,X})+O((T-T_{c,X})^2),
  \qquad
  A_X\neq0 .
\]
Third, the irrelevant microscopic coordinates determine finite-cutoff
corrections.  At linear order the tuning relation for a fixed thermal target
\(\lambda_{t,R}\) is
\[
  \lambda_t(\Lambda_0)
  =
  \left(\frac{\mu_R}{\Lambda_0}\right)^{y_t}
  \lambda_{t,R}
  +\hbox{higher-order terms},
\]
so
\[
  T-T_{c,X}
  =
  A_X^{-1}
  \left(\frac{\mu_R}{\Lambda_0}\right)^{y_t}
  \lambda_{t,R}
  +\cdots .
\]
The factor \(A_X\) is not fixed-point data.  The exponent
\(y_t=D-\Delta_\varepsilon\) is.

\begin{proposition}[Coordinate universality versus scaling-limit existence]
\label{prop:coordinate-universality-versus-scaling-limit-existence}
Proposition~\ref{prop:local-universality-rg-coordinates} is a statement about
convergence of finitely many projected RG coordinates at the reference scale
\(\mu_R\).  It does not, by itself, construct a continuum probability measure,
a Euclidean source functional, or a Wightman theory.  To upgrade the coordinate
statement to a continuum universality theorem one must additionally establish
the existence of the scaling limit of the relevant Schwinger functions or,
equivalently in a source-functional formulation, convergence of the
renormalized lattice source functionals on a source domain large enough to
determine the desired correlators.  The limiting Schwinger functions must also
satisfy the positivity, covariance, locality, regularity, and clustering
hypotheses needed for the intended reconstruction or structural theorem.
\end{proposition}

\begin{proof}
The local RG proposition controls the finite vector
\(z(\mu_R;\Lambda_0)\) after projecting the exact cutoff action to the chosen
coordinate chart.  A continuum measure or source functional contains
infinitely many correlation data and their positivity properties.  Convergence
of a finite coordinate vector does not imply tightness of the underlying
measures, convergence of all moments or distributions, reflection positivity
in the limit, locality of the reconstructed fields, or uniqueness of the
continuum state.  Each of these is an additional analytic assertion about the
regulated models.  Hence scaling-limit existence is an independent hypothesis
or theorem, not a conclusion of the coordinate estimate.
\end{proof}

\begin{hypothesis}[Continuum universality data]
\label{hyp:conditional-lattice-to-continuum-universality-data}
Let \(X\) range over a class of reflection-positive microscopic lattice
models with the same global symmetries as the Ising fixed point, and let
\(p_X(a,u_R)\) be microscopic parameters tuned so that the relevant endpoint
coordinate at the reference scale \(\mu_R\) is the same fixed value \(u_R\).
Assume the following hypotheses.
\begin{enumerate}[label=(\roman*)]
  \item The finite-coordinate RG hypotheses of
  Proposition~\ref{prop:local-universality-rg-coordinates} hold uniformly for
  the class \(X\), so that the retained Wilsonian coordinates satisfy
  \[
    z_X(a,u_R)\longrightarrow z_\ast(u_R):=(u_R,V_{\mu_R}(u_R)).
  \]
  \item For the selected microscopic observables and sources, there are local
  source renormalization maps \(\mathcal Z^X_a\) and source-local vacuum
  counterterms \(P^X_a\) on a common test-source neighborhood such that the
  renormalized source functionals have the representation
  \[
    W^X_a[\mathcal Z^X_a(\eta)]-W^X_a[0]-P^X_a(\eta)
    =
    \mathcal W_X(z_X(a,u_R),\eta)+r^X_a(\eta),
  \]
  where \(r^X_a(\eta)\to0\) as a distributional functional of \(\eta\).
  \item The continuum source coordinate has been fixed by a common
  normalization of the leading scaling fields, and the maps
  \(\mathcal W_X(z,\eta)\) are sequentially continuous at
  \(z=z_\ast(u_R)\) with the same limiting functional:
  \[
    \lim_{z\to z_\ast(u_R)}\mathcal W_X(z,\eta)=W_\ast(\eta;u_R)
  \]
  for every \(X\) in the class, modulo source-local polynomials.
  \item The reflection-positive lattice quadratic forms for the renormalized
  observables converge on the positive-time test-function domain, and the
  limiting source functional has the Euclidean covariance, locality,
  regularity, and clustering properties required for the intended
  Osterwalder--Schrader reconstruction.
\end{enumerate}
\end{hypothesis}

\begin{theorem}[Conditional lattice-to-continuum universality]
\label{thm:conditional-lattice-to-continuum-universality}
Under Hypothesis~\ref{hyp:conditional-lattice-to-continuum-universality-data},
the continuum separated-point Schwinger functions of the normalized
scaling fields are independent of \(X\).  More precisely, functional
derivatives of \(W_\ast(\eta;u_R)\) give a common distribution on the
configuration space of distinct insertion points, while the source-local
polynomials in
Hypothesis~\ref{hyp:conditional-lattice-to-continuum-universality-data}\textup{(iii)}
can change only contact terms on collision diagonals.  Under the positivity
and regularity hypotheses in
Hypothesis~\ref{hyp:conditional-lattice-to-continuum-universality-data}\textup{(iv)},
the common Schwinger functions reconstruct the same Euclidean-to-Lorentzian
QFT data in the
observable sector generated by these fields.  At \(u_R=0\), if the additional
stress-tensor hypothesis stated after
Hypothesis~\ref{hyp:ising-separated-spin-scaling-limit} holds, this common
fixed point is the corresponding Ising CFT.
\end{theorem}

\begin{proof}
By Hypothesis~\ref{hyp:conditional-lattice-to-continuum-universality-data}\textup{(i)},
all tuned microscopic realizations approach the same retained-coordinate point
\(z_\ast(u_R)\) at the reference scale.
Hypothesis~\ref{hyp:conditional-lattice-to-continuum-universality-data}\textup{(ii)}
separates the dependence of the renormalized source functional into a
coordinate-dependent part, a vanishing remainder, and source-local vacuum
terms.  Hypothesis~\ref{hyp:conditional-lattice-to-continuum-universality-data}\textup{(iii)}
then gives
\[
  W^X_a[\mathcal Z^X_a(\eta)]-W^X_a[0]-P^X_a(\eta)
  \longrightarrow
  W_\ast(\eta;u_R)
\]
for every \(X\), modulo source-local polynomials.  Taking functional
derivatives at \(\eta=0\) gives the limiting distributions of inserted
renormalized observables.  Source-local polynomial differences are supported
on collision diagonals, so they vanish after restriction to pairwise distinct
insertion points.  This proves equality of separated-point Schwinger
functions.  Hypothesis~\ref{hyp:conditional-lattice-to-continuum-universality-data}\textup{(iv)}
is precisely the additional closedness and
regularity input needed to pass from equality of separated kernels to a
common reconstructed Euclidean/Lorentzian theory; it is not supplied by the
finite-dimensional RG coordinate estimate.
\end{proof}

\begin{hypothesis}[Ising lattice-to-CFT universality data]
\label{hyp:conditional-ising-universality-lattice-to-cft-data}
Fix a spatial dimension \(D\) and a class \(\mathfrak U\) of
reflection-positive \(\mathbb Z_2\)-invariant microscopic lattice models whose
thermodynamic limits exist in the pure-state convention used above.  Let
\(A^X_{\sigma,a}\) and \(A^X_{\varepsilon,a}\) be odd and even microscopic
observables in a model \(X\in\mathfrak U\), normalized by nonzero constants
so that their leading scaling fields are \(\sigma\) and \(\varepsilon\).
Assume the following data and hypotheses.
\begin{enumerate}[label=(\roman*)]
  \item The tuned finite-coordinate Wilsonian hypotheses of
  Proposition~\ref{prop:local-universality-rg-coordinates} hold uniformly for
  \(\mathfrak U\), with relevant endpoint coordinate
  \(u_R=(\lambda_{h,R},\lambda_{t,R})\).  In the \(\mathbb Z_2\)-even critical
  slice this means \(\lambda_{h,R}=\lambda_{t,R}=0\).
  \item The source-dependent scaling-limit hypotheses of
  Theorem~\ref{thm:conditional-lattice-to-continuum-universality} hold for the
  source coordinates coupled to \(A^X_{\sigma,a}\), \(A^X_{\varepsilon,a}\),
  and the other selected scaling observables.  The reflected OS quadratic
  forms converge on the positive-time test-function domain.
  \item At \(u_R=0\), the common limiting Schwinger hierarchy
  \(W_\ast(\eta;0)\) satisfies Euclidean covariance, reflection positivity,
  regularity, clustering, and the radial reconstruction hypotheses of
  Hypothesis~\ref{hyp:radial-reconstruction-data}.  It has a conserved
  symmetric stress tensor improvable to traceless form, so the fixed point is
  a CFT in the sense of
  Chapter~\ref{ch:volume-iii-fixed-points-conformal-data}.
  \item The fixed-point local CFT data include the local scaling-operator
  spectrum, two-point pairings, OPE coefficients, stress-tensor Ward data, and
  the OPE convergence hypotheses of
  Definition~\ref{def:radial-ope-convergence-hypotheses}.  For nonzero
  \(u_R\), the corresponding relevant deformation of the fixed-point source
  functional by
  \[
    \int\dd^Dx\,
    \bigl(
      \lambda_{h,R}\mu_R^{D-\Delta_\sigma}\sigma(x)
      +
      \lambda_{t,R}\mu_R^{D-\Delta_\varepsilon}\varepsilon(x)
    \bigr)
  \]
  is constructed in a source-functional neighborhood large enough to determine
  the separated scaling functions under discussion.
\end{enumerate}
\end{hypothesis}

\begin{theorem}[Conditional Ising universality from lattice data to CFT data]
\label{thm:conditional-ising-universality-lattice-to-cft-data}
Under Hypothesis~\ref{hyp:conditional-ising-universality-lattice-to-cft-data},
for every \(X\in\mathfrak U\), every list of selected microscopic
observables, and every configuration of pairwise distinct continuum points
\(x_1,\ldots,x_n\), the normalized thermodynamic lattice correlators have the
common scaling limit
\[
  \lim_{a\downarrow0}
  \left\langle
    \prod_{j=1}^n
    a^{-\Delta_{I_j}}
    (b^X_{I_j})^{-1}A^X_{I_j,a}(x_j)
  \right\rangle_{X,p_X(a,u_R)}
  =
  S^{\rm Ising}_{I_1\ldots I_n}(x_1,\ldots,x_n;u_R),
  \label{eq:conditional-ising-universality-scaling-functions}
\]
where \(I_j\) labels \(\sigma,\varepsilon\), or another selected scaling
operator, \(b^X_{I_j}\neq0\) is the leading microscopic normalization, and
\(p_X(a,u_R)\) is the tuned microscopic parameter supplied by
Hypothesis~\ref{hyp:conditional-ising-universality-lattice-to-cft-data}\textup{(i)}.
At \(u_R=0\), the functions on the right are the separated-point
correlators of the Ising CFT.  They are determined by the fixed-point CFT
data in
Hypothesis~\ref{hyp:conditional-ising-universality-lattice-to-cft-data}\textup{(iv)}
through radial OPE convergence and the Ward identities.  At
nonzero \(u_R\) inside the constructed relevant-deformation neighborhood, the
same fixed-point data together with the specified relevant coordinates
determine the massive or magnetized scaling functions.

The quantities \(T_{c,X}\), metric factors relating \(T-T_{c,X}\) and the
magnetic field to \(\lambda_{t,R}\) and \(\lambda_{h,R}\), the constants
\(b^X_I\), finite-cutoff corrections, and contact-term extensions are not
part of the universal separated-point functions
\eqref{eq:conditional-ising-universality-scaling-functions}.
\end{theorem}

\begin{proof}
The thermodynamic limit supplies the infinite-volume lattice state for each
fixed cutoff and tuned microscopic parameter.
Hypothesis~\ref{hyp:conditional-ising-universality-lattice-to-cft-data}\textup{(i)},
by
Proposition~\ref{prop:local-universality-rg-coordinates}, sends every tuned
microscopic realization to the same retained Wilsonian coordinate
\[
  z_\ast(u_R)=(u_R,V_{\mu_R}(u_R))
\]
at the reference scale.
Hypothesis~\ref{hyp:conditional-ising-universality-lattice-to-cft-data}\textup{(ii)},
by
Theorem~\ref{thm:conditional-lattice-to-continuum-universality}, upgrades that
coordinate convergence to convergence of the renormalized source functionals:
after the local source maps and source-local vacuum subtractions are applied,
all \(X\in\mathfrak U\) have the same limiting functional
\(W_\ast(\eta;u_R)\) modulo source-local polynomials.  Functional derivatives
with respect to the normalized sources give
\eqref{eq:conditional-ising-universality-scaling-functions}.  Source-local
polynomials are supported on collision diagonals, so they vanish on the
configuration space \(x_i\neq x_j\).

At \(u_R=0\),
Hypothesis~\ref{hyp:conditional-ising-universality-lattice-to-cft-data}\textup{(iii)}
gives the OS/radial reconstruction data and the
stress-tensor condition needed to regard \(W_\ast(\eta;0)\) as the source
functional of a unitary CFT.
Hypothesis~\ref{hyp:conditional-ising-universality-lattice-to-cft-data}\textup{(iv)}
supplies the local CFT data and
OPE convergence.  The radial OPE theorem then reconstructs separated-point
correlators from the local scaling spectrum, two-point pairings, OPE
coefficients, and Ward data, with the contact chart fixed by the source
functional.  For \(u_R\neq0\) in the stated source-functional neighborhood,
the deformation by \(\sigma\) and \(\varepsilon\) is part of the constructed
functional data, so the same argument gives the corresponding scaling
functions with the relevant coordinates held fixed.  The last sentence lists
parameters removed either by tuning, by field normalization, or by restriction
to separated points; the preceding argument never identifies them with CFT
data.
\end{proof}

The same separation applies to observables.  Let \(A^X_{\rm spin}\) be a
microscopic odd lattice observable and \(A^X_{\rm eng}\) a microscopic even
energy-density observable with identity part subtracted.  Their scaling-field
expansions have the form
\[
  A^X_{\rm spin}
  =
  b^X_\sigma a^{\Delta_\sigma}\sigma
  +
  \sum_J b^X_J a^{\Delta_J}O_J,
  \qquad
  A^X_{\rm eng}
  =
  b^X_\varepsilon a^{\Delta_\varepsilon}\varepsilon
  +
  \sum_K c^X_K a^{\Delta_K}O_K,
\]
where the sums contain fields with larger scaling dimensions in the same
symmetry sector, and \(b^X_\sigma,b^X_\varepsilon\) are nonzero for the
chosen observables.

Assume now, in addition to the coordinate hypotheses above, that the
renormalized separated-point lattice correlators for the selected observables
have scaling limits as distributions and that the coordinate-to-correlator map
is continuous at the limiting graph \(v=V_{\mu_R}(u_R)\).  For separated
continuum points \(x_i\), the normalized lattice correlators then have the
common limit
\[
  \lim_{a\to0}
  \prod_i
  \left((b^X_i)^{-1}a^{-\Delta_i}\right)
  \left\langle
    \prod_i A^X_i([x_i/a])
  \right\rangle_{X,a,p_X(a,u_R)}
  =
  \left\langle
    \prod_i O_i(x_i)
  \right\rangle_{u_R},
\]
with the same limiting continuum correlator for all \(X\) satisfying the
coordinate hypotheses and the additional scaling-limit hypotheses.  The
constants \(b^X_i\), the critical temperature \(T_{c,X}\), and finite-cutoff
corrections are microscopic.  The dimensions, RG eigenvalues, and normalized
continuum correlators are fixed-point data once the continuum limit has been
constructed.

If the coordinate-to-correlator map is differentiable in the retained
irrelevant coordinates, the same estimate gives the standard correction
structure with all symbols fixed.  Let
\[
  G^X_a(x_1,\ldots,x_n)
\]
denote a normalized separated-point lattice correlator after solving the
relevant endpoint equation for \(u_R\).  In the finite chart,
\[
  G^X_a
  =
  G_{u_R}
  +
  \sum_\rho
  \left(
    v^\rho_X(\mu_R;\Lambda_0)-V^\rho_{\mu_R}(u_R)
  \right)
  \left.
    \frac{\partial G}{\partial v^\rho}
  \right|_{v=V_{\mu_R}(u_R)}
  +
  \cdots .
\]
For bounded microscopic irrelevant coordinates this gives
\[
  G^X_a(x_1,\ldots,x_n)
  =
  G_{u_R}(x_1,\ldots,x_n)
  +
  O((\mu_R a)^\kappa),
  \qquad
  a=\Lambda_0^{-1}.
\]
When the generated-integral convergence is faster than the leading irrelevant
correction, the smallest irrelevant exponent in the symmetry sector gives the
leading correction-to-scaling exponent.  The amplitudes multiplying these
corrections depend on \(X\); the exponents and the limiting correlator do not,
under the hypotheses stated above.
