\chapter{The Wilson-Fisher Fixed Point and Scaling Operators}
\label{ch:volume-ii-wilson-fisher-fixed-point-scaling-operators}
\ConventionsEuclidean


\section{Scalar Quartic Theory Below Four Dimensions}

The trace identity makes fixed points structurally important.  The next task is
to construct the perturbative data of a non-Gaussian fixed point in a
controlled small-\(\epsilon\) calculation.  Consider a single real scalar field in
\[
  D=4-\epsilon
\]
Euclidean dimensions.

\begin{definition}[Dimensional-regularized perturbative status]
\label{def:wilson-fisher-dimensional-regularization-status}
The calculation in this chapter is a dimensionally regularized perturbative
calculation.  At each fixed loop order and for each fixed graph, the symbol
\[
  \int \frac{\dd^{\,4-\epsilon}\ell}{(2\pi)^{4-\epsilon}}\,
  F(\ell,p,m)
\]
denotes the meromorphic continuation in the complex variable
\(D=4-\epsilon\) of the corresponding rotationally invariant Euclidean
loop-momentum integral, defined first in a domain where the graph integral or
its Feynman-parameter representation converges.  Subtractions are then applied
coefficient by coefficient in the loop expansion.  Thus the graph rules use
analytic continuation of finite-dimensional momentum integrals; no
field-configuration-space measure in noninteger dimension is being
constructed.

The quantities
\(\beta^\epsilon(\lambda)\), \(\gamma_\phi(\lambda)\),
\(\gamma_2(\lambda)\), \(\lambda_*(\epsilon)\), and the scaling dimensions
extracted from them are formal perturbative objects in a specified
subtraction scheme.  Before solving for the fixed point, the beta function and
anomalous dimensions are formal power series in the dimensionless coupling
\(\lambda\), with coefficients meromorphic, and after subtraction regular, as
\(\epsilon\to0\) in the displayed orders.  The Wilson--Fisher fixed-point
coordinate is obtained as a formal small-\(\epsilon\) branch
\(\lambda_*(\epsilon)\in\epsilon\,\mathbb R[[\epsilon]]\).  Substituting this
branch into anomalous dimensions gives formal \(\epsilon\)-expansions for
\(\Delta_\phi\), \(\Delta_{\phi^2}\), and other scaling data.  Convergence or
summability of these formal series is an additional analytic question.
\end{definition}

The local action density used to generate the perturbative graph rules is
\[
  \mathcal L_E
  =
  \frac12\partial_\mu\phi\,\partial^\mu\phi
  +
  \frac12m_0^2\phi^2
  +
  \frac1{4!}g_0\phi^4 .
\]
This action is invariant under the \(\mathbb Z_2\) transformation
\(\phi\mapsto-\phi\).  Unless an odd source is explicitly introduced, the
renormalization-group chart used in this chapter is the \(\mathbb Z_2\)-even
one.  This convention matters for the count of relevant tuning conditions:
inside the even chart the mass perturbation is the relevant deformation that
must be tuned, while in the full local theory space an odd source for
\(\phi\) is also a relevant coordinate.
The quartic coupling \(g_0\) has mass dimension \(\epsilon\).  The mass
parameter \(m_0^2\) must be tuned so that the infrared two-point function has
no mass gap.  This tuning is an infrared condition: it is not guaranteed by
the algebra of counterterms, and it is meaningful only when the flow reaches a
massless long-distance theory.  The symbol \(m_0^2\) here denotes the bare
quadratic coordinate in the cutoff action, not the spectral pole mass of
Definition~\ref{def:mass-data-and-mass-coordinates}.  The massless
Wilson--Fisher point is obtained by tuning this coordinate to the critical
surface; massive departures from the fixed point are measured by a
renormalized relevant coordinate, and only after reconstruction and spectral
analysis can that coordinate be related to a physical pole or gap.
The dimensionless quartic coordinate used below is therefore not the
dimensionful coefficient \(g_0\).  In the notation of the minimal-subtraction
chapter, \(g_0\) is the bare \(D=4-\epsilon\) coupling \(g^\epsilon\), and a
renormalized quartic coordinate is obtained by factoring out the engineering
dimension:
\[
  g_0
  =
  \mu^\epsilon
  \left[
    \lambda(\mu)+\Delta_\lambda(\lambda(\mu),\epsilon)
  \right],
  \qquad
  g_R(\mu):=\mu^\epsilon\lambda(\mu),
  \qquad
  \Delta_\lambda=O(\lambda^2).
\]
The symbol \(\lambda(\mu)\) in the momentum-subtraction calculation immediately
below denotes the dimensionless coordinate.  The term \(-\epsilon\lambda\) in
the beta function is precisely the derivative of the factor
\(\mu^{-\epsilon}g_0\) at fixed bare coupling.

\begin{definition}[Symmetric MOM quartic chart]
\label{def:wilson-fisher-symmetric-mom-chart}
In the one-component \(\mathbb Z_2\)-even scalar theory, the symmetric
Euclidean momentum-subtraction chart at scale \(\mu>0\) is the local
renormalized coordinate system defined as follows.  The dimensionless quartic
coordinate is the 1PI coefficient
\[
  \lambda(\mu)
  =
  -\mu^{-\epsilon}Z(\mu)^2\,
  \Gamma^{(4)}
  \bigg|_{k_i^2=\mu^2,\;s=t=u=-\frac43\mu^2},
\]
where
\[
  Z(\mu)
  =
  \left(
    1-\frac{\partial\Sigma}{\partial k^2}
  \right)^{-1}_{k^2=\mu^2}.
\]
The subtraction point is nonexceptional: no nonempty proper subset of the
external momenta has vanishing total momentum.  The mass coordinate of this
chart is fixed separately by the endpoint condition that the infrared
two-point function be massless.  Thus the displayed \(\lambda(\mu)\) is a
coordinate on the massless endpoint slice; the mass tuning is the separate
condition that places the trajectory on that slice.

If several schemes are present, the coordinates of this chart will be denoted
\(\lambda_{\rm MOM}(\mu)\) and \(Z_{\rm MOM}(\mu)\); the lighter notation
\(\lambda(\mu)\) and \(Z(\mu)\) is used here only while no pole-subtraction
field factor is being discussed.  In particular, \(Z_{\rm MOM}(\mu)\) is a
finite residue normalization at the subtraction momentum, whereas the
minimal-subtraction field factor introduced below contains only pole
counterterms in \(\epsilon\).
\end{definition}

\begin{proposition}[One-loop quartic vector field]
\label{prop:wilson-fisher-one-loop-quartic-vector-field}
In the chart of
Definition~\ref{def:wilson-fisher-symmetric-mom-chart}, the one-loop
dimensionally regularized quartic RG vector field is
\[
  \beta^\epsilon(\lambda)
  =
  \frac{\dd\lambda}{\dd\log\mu}
  =
  -\epsilon\lambda
  +
  \left(\frac{3}{16\pi^2}+O(\epsilon)\right)\lambda^2
  +
  O(\lambda^3).
\]
Consequently the nonzero perturbative zero in the one-coupling massless chart
has the formal small-\(\epsilon\) expansion
\[
  \lambda_*(\epsilon)
  =
  \frac{16\pi^2}{3}\epsilon+O(\epsilon^2).
\]
\end{proposition}

\begin{proof}
At one loop \(Z(\mu)=1\), and the bubble graph gives
\[
  \lambda(\mu)
  =
  \mu^{-\epsilon}
  \left[
    g_0
    -
    \frac32g_0^2(4\pi)^{-2+\epsilon/2}
    \frac{\Gamma(\epsilon/2)\Gamma(1-\epsilon/2)^2}{\Gamma(2-\epsilon)}
    \left(\frac43\mu^2\right)^{-\epsilon/2}
    +
    O(g_0^3)
  \right].
\]
Let
\[
  A_0(\epsilon)
  =
  -\frac32(4\pi)^{-2+\epsilon/2}
  \frac{\Gamma(\epsilon/2)\Gamma(1-\epsilon/2)^2}{\Gamma(2-\epsilon)}
  \left(\frac43\right)^{-\epsilon/2}.
\]
Then the preceding equation has the form
\[
  \lambda(\mu)
  =
  \mu^{-\epsilon}g_0
  +
  A_0(\epsilon)\,\mu^{-2\epsilon}g_0^2
  +
  O(g_0^3).
\]
Differentiating at fixed bare coupling gives
\[
  \frac{\dd\lambda}{\dd\log\mu}
  =
  -\epsilon\,\mu^{-\epsilon}g_0
  -
  2\epsilon\,A_0(\epsilon)\,\mu^{-2\epsilon}g_0^2
  +
  O(g_0^3).
\]
To express the right-hand side in terms of \(\lambda(\mu)\), use
\[
  \mu^{-\epsilon}g_0
  =
  \lambda(\mu)-A_0(\epsilon)\lambda(\mu)^2+O(\lambda^3),
  \qquad
  \mu^{-2\epsilon}g_0^2
  =
  \lambda(\mu)^2+O(\lambda^3).
\]
Thus
\[
  \frac{\dd\lambda}{\dd\log\mu}
  =
  -\epsilon\lambda
  -
  \epsilon\lambda^2
  \left[
    -\frac32(4\pi)^{-2+\epsilon/2}
    \frac{\Gamma(\epsilon/2)\Gamma(1-\epsilon/2)^2}{\Gamma(2-\epsilon)}
    \left(\frac43\right)^{-\epsilon/2}
  \right]
  +
  O(\lambda^3).
\]
The pole in the bracket is multiplied by the engineering factor \(\epsilon\);
the result is finite order by order in the small but nonzero parameter
\(\epsilon\).
Since
\[
  -\frac32(4\pi)^{-2+\epsilon/2}
  \frac{\Gamma(\epsilon/2)\Gamma(1-\epsilon/2)^2}{\Gamma(2-\epsilon)}
  \left(\frac43\right)^{-\epsilon/2}
  =
  -\frac{3}{16\pi^2}\frac1{\epsilon}+O(\epsilon^0),
\]
the \(D=4-\epsilon\) beta function is denoted \(\beta^\epsilon\):
\[
  \beta^\epsilon(\lambda)
  =
  \frac{\dd\lambda}{\dd\log\mu}
  =
  -\epsilon\lambda
  +
  \left(\frac{3}{16\pi^2}+O(\epsilon)\right)\lambda^2
  +
  O(\lambda^3).
\]
This notation distinguishes the \(4-\epsilon\)-dimensional vector field from
the strictly four-dimensional quartic beta function discussed in the 1PI RG
chapter, which is obtained by setting the engineering term to zero.  The zero
of \(\beta^\epsilon\) is
\[
  \lambda_*
  =
  \frac{16\pi^2}{3}\epsilon+O(\epsilon^2).
\]
For sufficiently small \(\epsilon\), the formal fixed-point coordinate lies in
the perturbative domain in the following precise sense.  The calculation
constructs a formal power series
\[
  \lambda_*(\epsilon)
  =
  \sum_{r\ge1}a_r\epsilon^r,
  \qquad
  a_1=\frac{16\pi^2}{3},
\]
whose coefficients are determined recursively from the beta function expanded
as a formal series in \(\lambda\) and \(\epsilon\).  At any finite loop order,
one obtains a truncated polynomial vector field and a zero whose residual in
the next omitted order is of higher order in \(\epsilon\).  An exact RG flow
or a nonperturbative continuum QFT with a fixed point at fixed positive
\(\epsilon\) requires additional input: for example a constructive Wilsonian
or lattice continuum-limit theorem, or a summability theorem for the
perturbative data together with a proof that the summed correlators satisfy
the required locality, positivity, and regularity properties.  In this
chapter, unless a separate nonperturbative construction is explicitly stated,
``the Wilson--Fisher fixed point'' means this formal perturbative fixed-point
data.  The mass perturbation is a separate relevant direction and will be
identified below from the \(\phi^2\) scaling operator.
\end{proof}

\begin{remark}[Local uniqueness of the perturbative branch]
\label{rem:wilson-fisher-local-uniqueness}
The uniqueness statement supplied by this calculation is local and
chart-dependent.  Restrict to the one-component \(\mathbb Z_2\)-even scalar
chart in which the mass has been tuned and the only displayed dimensionless
interaction coordinate is \(\lambda\).  Put
\[
  \beta^\epsilon(\lambda)
  =
  -\epsilon\lambda+b\lambda^2+O(\epsilon\lambda^2,\lambda^3),
  \qquad
  b=\frac{3}{16\pi^2}.
\]
For a nonzero perturbative zero write \(\lambda=\epsilon u\).  Then
\[
  \epsilon^{-2}\beta^\epsilon(\epsilon u)
  =
  -u+bu^2+O(\epsilon).
\]
At \(\epsilon=0\) the nonzero root is
\[
  u_0=b^{-1}=\frac{16\pi^2}{3},
\]
and
\[
  \frac{\partial}{\partial u}(-u+bu^2)\bigg|_{u=u_0}=1\neq0 .
\]
Equivalently, solving order by order in \(\epsilon\) determines a unique
formal branch \(u(\epsilon)=u_0+O(\epsilon)\) in a neighborhood of \(u_0\).
Together with the Gaussian branch \(\lambda=0\), this exhausts the small
one-coupling branches visible in this local chart.

The scope is local uniqueness in the one-component \(\mathbb Z_2\)-even
small-coupling scalar chart.  Other perturbative fixed-point constructions
belong to different charts or field contents: \(O(N)\)-vector
Wilson--Fisher fixed points, multicritical scalar fixed points such as the
tricritical \(\phi^6\) branch near \(D=3-\epsilon\), multi-coupling gauge or
gauge--Yukawa fixed points of Banks--Zaks type, and decoupled product fixed
points.  Nonperturbative fixed points outside the small-coupling neighborhood
require separate construction.
\end{remark}

Figure~\ref{fig:volume-ii-wilson-fisher-flow} displays the local
one-coordinate Wilson--Fisher construction: the symmetric Euclidean 1PI quartic
vertex defines the coordinate, the beta function has the perturbative fixed
point, and the critical trajectory is obtained after imposing the mass endpoint
condition.

\begin{figure}[H]
  \centering
  \resizebox{0.98\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small, align=center] at (0,2.45)
        {quartic coordinate from a symmetric\\Euclidean 1PI vertex};
      \fill[gray!18, draw=black, thick] (0,0.45) circle (0.52);
      \node at (0,0.45) {1PI};
      \foreach \a/\lab in {35/\(k_1\),145/\(k_2\),215/\(k_3\),325/\(k_4\)} {
        \draw[thick] ({0.52*cos(\a)},{0.45+0.52*sin(\a)})
          --++({1.05*cos(\a)},{1.05*sin(\a)});
        \node at ({1.85*cos(\a)},{0.45+1.85*sin(\a)}) {\lab};
      }
      \node[align=center, blue!70!black] at (0,-1.25)
        {\(\lambda(\mu)=-\mu^{-\epsilon}Z(\mu)^2\Gamma^{(4)}\)\\
         \(k_i^2=\mu^2,\quad s=t=u=-4\mu^2/3\)};
    \end{scope}

    \begin{scope}[shift={(6.7,0)}]
      \node[font=\small, align=center] at (0,2.45)
        {flow for decreasing \(\mu\)};
      \draw[-{Latex[length=2mm]}, thick] (-3.0,0)--(3.45,0) node[below] {\(\lambda\)};
      \fill[black] (-2.25,0) circle (2pt);
      \node[below] at (-2.25,-0.08) {\(0\)};
      \fill[purple!80!black] (0.35,0) circle (2.5pt);
      \node[purple!80!black, below] at (0.35,-0.08) {\(\lambda_*\)};
      \foreach \x in {-1.75,-1.05,-0.35} {
        \draw[-{Latex[length=2mm]}, red!75!black, very thick] (\x,0.32)--++(0.52,0);
      }
      \foreach \x in {1.45,2.15,2.85} {
        \draw[-{Latex[length=2mm]}, blue!75!black, very thick] (\x,0.32)--++(-0.52,0);
      }
      \node[red!75!black] at (-1.05,1.05) {\(\beta^\epsilon<0\)};
      \node[blue!75!black] at (2.15,1.05) {\(\beta^\epsilon>0\)};
      \node[purple!80!black, align=center] at (0.35,-1.05)
        {formal IR-attractive\\on the massless endpoint set};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The dimensionless quartic coordinate is defined by a symmetric 1PI
  subtraction.  In the formal \(4-\epsilon\) perturbative vector field the
  classical term \(-\epsilon\lambda\) and the one-loop term
  \(3\lambda^2/(16\pi^2)\) balance at the nonzero zero \(\lambda_*\).  The
  displayed one-dimensional flow is the perturbative flow after the mass
  endpoint condition has been imposed.}
  \label{fig:volume-ii-wilson-fisher-flow}
\end{figure}

The bare coupling and the renormalized coupling have different infrared
behavior.  The dimensionless bare coordinate \(\mu^{-\epsilon}g_0\)
diverges as \(\mu\to0\).  In the formal perturbative RG equation, the
renormalized coupling \(\lambda(\mu)\), after the fluctuation correction in
the 1PI coordinate definition is included, has the endpoint value
\(\lambda_*(\epsilon)\) along the massless critical trajectory.

Figure~\ref{fig:volume-ii-bare-versus-renormalized-coupling} separates these two
coordinate behaviors: the dimensionless bare coordinate diverges under the
infrared rescaling, while the renormalized coordinate approaches the
Wilson--Fisher fixed-point value.

\begin{figure}[H]
  \centering
  \resizebox{0.85\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \draw[-{Latex[length=2mm]}, thick] (-3.2,0)--(3.5,0) node[below] {\(\mu\downarrow0\)};
    \node[align=center] at (-1.8,1.78)
      {bare coordinate\\\(\mu^{-\epsilon}g_0\to\infty\)};
    \draw[red!75!black, very thick, -{Latex[length=2mm]}]
      (-2.55,0.32)..controls (-1.8,0.72) and (-1.05,1.02)..(-0.2,1.28);
    \node[align=center] at (1.55,1.45)
      {1PI coordinate\\formal endpoint \(\lambda_*\)};
    \draw[blue!75!black, very thick, -{Latex[length=2mm]}]
      (0.15,0.35)..controls (0.95,0.72) and (1.75,0.78)..(2.55,0.78);
    \draw[purple!80!black, dashed] (2.55,0.05)--(2.55,1.0);
    \node[purple!80!black, right] at (2.55,0.78) {\(\lambda_*\)};
  \end{tikzpicture}
  }
  \caption{The finite infrared coupling in the formal perturbative flow is
  defined by the renormalized 1PI coordinate.  The dimensionless bare
  coordinate grows without approaching a finite fixed-point value.}
  \label{fig:volume-ii-bare-versus-renormalized-coupling}
\end{figure}

\section{The First Scaling Operators}

\begin{definition}[Scaling operators in a local source chart]
\label{def:wilson-fisher-scaling-operator-source-chart}
Fix a renormalized local source chart near a fixed point.  Thus one has a
finite set of renormalized local fields \(O_A\), with total derivatives,
equation-of-motion fields, and the identity source either removed or placed in
separate blocks, and a source deformation
\[
  \delta S
  =
  \sum_A\int \dd^D x\, g^A(x)\,O_A(x).
\]
For constant sources, let \(g^A\) denote the corresponding dimensionless
coordinates and suppose that their RG equation has a linearization
\[
  \frac{\dd g^A}{\dd\log\mu}
  =
  -M^A{}_B\,g^B+O(g^2)
\]
at the fixed point.  A scaling source is an eigen-coordinate \(u^r\) of
\(M\).  If
\[
  \frac{\dd u^r}{\dd\log\mu}
  =
  -y_r u^r+O(g^2),
\]
then the dual scaling operator has dimension
\[
  \Delta_r=D-y_r .
\]
Equivalently, under a change of physical scale \(\mu\mapsto s\mu\) at the
linearized fixed point, \(u^r\) scales as \(s^{-y_r}u^r\).  Relevant,
marginal, and irrelevant directions are respectively those with
\(y_r>0\), \(y_r=0\), and \(y_r<0\).  This definition is local in the chosen
source chart; if several fields with the same quantum numbers mix, the
operators themselves are the eigenvectors of the mixing matrix, not the
individual basis representatives.
\end{definition}

The elementary field \(O_1=\phi\) has a fixed-point scaling dimension
independent of finite analytic changes of the local field coordinate, but the
anomalous-dimension function away from the fixed point depends on the
renormalization chart.  We compute it below in the minimal-subtraction pole
chart.  Let \(\lambda_{\rm MS}\) be the dimensionless quartic coordinate in
this chart, and let
\[
  \phi_0=\left(Z_\phi^{\rm MS}\right)^{1/2}\phi_{\rm MS},
  \qquad
  Z_\phi^{\rm MS}(\lambda_{\rm MS},\epsilon)
  =
  1+\sum_{r\geq 1}\frac{Z_{\phi,r}(\lambda_{\rm MS})}{\epsilon^r},
\]
where \(\phi_0\) is the regulated field appearing in the bare graph expansion
and \(Z_\phi^{\rm MS}\) contains no finite part.  The field anomalous dimension
in this chart is
\[
  \gamma_\phi^{\rm MS}(\lambda_{\rm MS})
  =
  \left.
  \frac12\mu\frac{\dd}{\dd\mu}\log Z_\phi^{\rm MS}
  \right|_{g_0,\epsilon}.
\]
Since \(Z_\phi^{\rm MS}=1+O(\lambda_{\rm MS}^2)\) in scalar quartic theory, the
field anomalous dimension starts only at order
\(\lambda_{{\rm MS},*}^2=O(\epsilon^2)\).  Thus
\[
  \Delta_\phi
  =
  \frac{D-2}{2}+\gamma_{\phi*}^{\rm MS}
  =
  1-\frac{\epsilon}{2}+O(\epsilon^2).
\]
The conventional critical exponent \(\eta\) is defined by the long-distance
two-point decay
\[
  \left\langle\phi(x)\phi(0)\right\rangle
  \sim
  |x|^{-D+2-\eta}.
\]
Since a scalar primary of dimension \(\Delta_\phi\) has two-point exponent
\(2\Delta_\phi\), this convention gives
\[
  \eta=2\Delta_\phi-(D-2)=2\gamma_{\phi *}^{\rm MS}.
\]
The first nonzero field anomalous dimension comes from the momentum-dependent
part of the two-loop sunset self-energy.  In the following pole computation
write \(\lambda\) for \(\lambda_{\rm MS}\).  With the interaction normalized as
\(\lambda\phi^4/4!\), the sunset graph has symmetry factor \(1/6\), hence
\[
  \Sigma_{\rm sun}^{(2)}(k)
  =
  \frac{\lambda^2}{6}\,\mu^{2\epsilon} I(k),
  \qquad
  I(k)=
  \int_{p,q}
  \frac{1}{p^2q^2(p+q+k)^2},
  \qquad
  \int_p:=\int\frac{\dd^{4-\epsilon}p}{(2\pi)^{4-\epsilon}} .
\]
First integrate over \(q\):
\[
  \int_q\frac{1}{q^2(q+p+k)^2}
  =
  \frac{1}{(4\pi)^{D/2}}
  \frac{\Gamma(\epsilon/2)\Gamma(1-\epsilon/2)^2}{\Gamma(2-\epsilon)}
  \bigl((p+k)^2\bigr)^{-\epsilon/2},
  \qquad D=4-\epsilon .
\]
The remaining massless two-point integral gives
\[
\begin{aligned}
  I(k)
  &=
  \frac{(k^2)^{1-\epsilon}}{(4\pi)^D}
  \frac{\Gamma(1-\epsilon/2)^3\Gamma(-1+\epsilon)}
       {\Gamma(3-3\epsilon/2)}
  \\
  &=
  -\,\frac{k^2}{2(4\pi)^4}\,\frac1\epsilon
  + \hbox{finite and higher-order terms in }\epsilon .
\end{aligned}
\]
The factor \(1/2\) is the \(\Gamma(3)^{-1}\) value in this last two-point
integral, equivalently the corresponding Feynman-parameter normalization.
Multiplying it by the sunset symmetry factor \(1/6\) gives the coefficient
\(1/12\).  Therefore the divergent part of the derivative is
\[
  \left.
  \frac{\partial\Sigma^{(2)}(k)}{\partial k^2}
  \right|_{\rm div}
  =
  -\frac{\lambda^2}{12(16\pi^2)^2}\,\frac1\epsilon ,
\]
where the self-energy sign convention is the same as in the
momentum-subtraction chart, but the subtraction now retains only the pole
part.  Write
\[
  a(\lambda)=\frac{\lambda^2}{12(16\pi^2)^2}.
\]
The pole part of the minimal-subtraction field factor is determined by the same
residue condition:
\[
  Z_\phi^{\rm MS}
  =
  \left[
    \left(1-\partial_{k^2}\Sigma\right)^{-1}
  \right]_{\rm pole}
  =
  1-\frac{a(\lambda)}{\epsilon}
  +O(\lambda^3,\epsilon^0)
\]
where \([\cdot]_{\rm pole}\) means that all finite terms are discarded.  The
sign in the anomalous dimension is then fixed by differentiating at fixed
regulator definition.  To the order needed here
\(\beta_{\rm MS}(\lambda)=-\epsilon\lambda+O(\lambda^2)\), so
\[
  \frac12\mu\frac{\dd}{\dd\mu}\log Z_\phi^{\rm MS}
  =
  \frac12\beta_{\rm MS}(\lambda)\frac{\partial}{\partial\lambda}
  \left(-\frac{a(\lambda)}{\epsilon}\right)
  =
  a(\lambda)+O(\lambda^3).
\]
Therefore
\[
  \gamma_\phi^{\rm MS}(\lambda)
  =
  \frac{\lambda^2}{12(16\pi^2)^2}+O(\lambda^3).
\]
Substituting \(\lambda_{{\rm MS},*}=(16\pi^2/3)\epsilon+O(\epsilon^2)\) gives
\[
  \eta=\frac{\epsilon^2}{54}+O(\epsilon^3)
\]
in the interaction normalization \(\lambda\phi^4/4!\).
This value is not an artifact of the pole chart.  If a second chart is obtained
by a finite analytic coupling redefinition
\(\lambda'=\lambda+c_2\lambda^2+O(\lambda^3)\) and the finite field-coordinate
change \(\phi'=F(\lambda)^{-1}\phi\), \(F(0)=1\), then
\[
  \gamma'_\phi(\lambda')
  =
  \gamma_\phi(\lambda)
  +
  \beta(\lambda)\frac{\partial\log F(\lambda)}{\partial\lambda}.
\]
At an exact fixed point the second term vanishes.  In the present
\(\epsilon\)-expansion, replacing \(\lambda_*\) by
\(\lambda'_*= \lambda_*+O(\lambda_*^2)\) changes
\(\gamma_\phi=A\lambda_*^2+O(\lambda_*^3)\) only at order \(O(\epsilon^3)\).

The operator \(O_2=\phi^2\) is more delicate.  The constant source deformation
\[
  \delta S=\int\dd^D x\,\delta g_2\,\phi^2(x)
\]
corresponds to a zero-momentum insertion.  In the massless theory, the
one-loop graph then contains an infrared-divergent integral of the form
\[
  \int\frac{\dd^D\ell}{(2\pi)^D}\,\frac{1}{(\ell^2)^2}.
\]
The divergence is caused by the subtraction kinematics, not by the local UV
definition of the operator.  Equivalently, the scaling-degree extension of the
coincident insertion controls the local counterterm ambiguity only after the
massless infrared prescription has supplied a separated-point distribution to
extend.

Use instead a spacetime-dependent source:
\[
  \delta S
  =
  \int\dd^D x\,\delta g_2(x)\,O_2(x)
  =
  \int\frac{\dd^Dp}{(2\pi)^D}\,
  \delta\widetilde g_2(p)\,\widetilde O_2(-p).
\]
Define the renormalized source by a nonzero-momentum 1PI insertion,
\[
  \delta\widetilde g_2(p;\mu)
  =
  -Z(\mu)
  \left(\text{amputated 1PI insertion of }O_2\right)
  \bigg|_{k^2=(p+k)^2=\mu^2}.
\]
The renormalized operator is determined by preserving the deformation:
\[
  \delta\widetilde g_2(p)\,\widetilde O_2(-p)
  =
  \delta\widetilde g_2(p;\mu)\,[\widetilde O_2(-p)]_\mu.
\]
At \(p^2=\mu^2\), define
\[
  \gamma_2(\mu)
  =
  \frac{\dd}{\dd\log\mu}
  \log
  \left(
    \frac{\delta\widetilde g_2(p;\mu)}
         {\delta\widetilde g_2(p)}
  \right).
\]

Figure~\ref{fig:volume-ii-phi-two-renormalization} identifies the two insertion
normalizations entering \(\gamma_2\): the zero-momentum composite insertion and
the fixed-momentum subtraction used to define the renormalized \(\phi^2\)
coordinate.

\begin{figure}[H]
  \centering
  \resizebox{0.98\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small, align=center] at (0,2.55)
        {zero-momentum insertion};
      \draw[thick] (-1.85,0.7)--(1.85,0.7);
      \fill[red!75!black] (0,0.7) circle (2.6pt);
      \draw[thick] (-0.75,0.7)..controls (-0.45,1.35) and (0.45,1.35)..(0.75,0.7);
      \draw[thick] (-0.75,0.7)..controls (-0.45,0.05) and (0.45,0.05)..(0.75,0.7);
      \node[red!75!black, above] at (0,1.05) {\(\delta g_2\)};
      \node[align=center, red!75!black] at (0,-0.8)
        {\(\displaystyle\int\frac{\dd^D\ell}{(\ell^2)^2}\)\\IR divergent};
    \end{scope}

    \begin{scope}[shift={(5.0,0)}]
      \node[font=\small, align=center] at (0,2.55)
        {momentum-carrying source};
      \fill[gray!18, draw=black, thick] (0,0.7) circle (0.48);
      \node at (0,0.7) {1PI};
      \draw[thick] (-2.05,0.1)--(-0.42,0.58);
      \draw[thick] (-2.05,1.3)--(-0.42,0.82);
      \draw[thick] (0.48,0.7)--(1.95,0.7);
      \fill[red!75!black] (1.95,0.7) circle (2.5pt);
      \node[below] at (-2.05,0.08) {\(k\)};
      \node[above] at (-2.05,1.32) {\(p+k\)};
      \node[red!75!black, above] at (1.95,1.0) {\(p\)};
      \node[align=center, blue!70!black] at (0,-0.85)
        {\(k^2=(p+k)^2=\mu^2\)};
    \end{scope}

    \begin{scope}[shift={(10.1,0)}]
      \node[font=\small, align=center] at (0,2.55)
        {one-loop anomalous\\dimension};
      \draw[thick] (-1.85,0.15)--(-0.82,0.55);
      \draw[thick] (-1.85,1.25)--(-0.82,0.85);
      \fill[red!75!black] (0,0.7) circle (2.6pt);
      \draw[thick] (-0.82,0.55)..controls (-0.45,1.35) and (0.45,1.35)..(0.82,0.85);
      \draw[thick] (-0.82,0.55)..controls (-0.45,0.05) and (0.45,0.05)..(0.82,0.85);
      \draw[thick] (0.82,0.85)--(1.85,1.2);
      \draw[thick] (0.82,0.55)--(1.85,0.2);
      \node[blue!70!black, align=center] at (0,-0.85)
        {\(\gamma_2=\lambda/(16\pi^2)+\cdots\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The \(\phi^2\) operator is renormalized by coupling it to a source
  with nonzero momentum.  This avoids the infrared singularity of the
  zero-momentum subtraction and gives a finite definition of the anomalous
  dimension.}
  \label{fig:volume-ii-phi-two-renormalization}
\end{figure}

The one-loop source relation is
\[
  \delta\widetilde g_2(p;\mu)
  =
  \delta\widetilde g_2(p)
  \left[
    1
    -
    \frac{g_4}{2}
    \int
    \frac{\dd^{4-\epsilon}\ell}{(2\pi)^{4-\epsilon}}
    \frac{1}{\ell^2(p+\ell)^2}
    +
    O(g_4^2)
  \right],
\]
or
\[
  \delta\widetilde g_2(p;\mu)
  =
  \delta\widetilde g_2(p)
  \left[
    1
    -
    \frac{g_4}{2}
    (4\pi)^{-2+\epsilon/2}
    \frac{\Gamma(\epsilon/2)\Gamma(1-\epsilon/2)^2}{\Gamma(2-\epsilon)}
    |p|^{-\epsilon}
    +
    O(g_4^2)
  \right].
\]
It follows that
\[
  \gamma_2(\mu)
  =
  \lambda(\mu)
  \left(
    \frac{1}{16\pi^2}+O(\epsilon)
  \right)
  +
  O(\lambda(\mu)^2).
\]
At the Wilson-Fisher fixed point,
\[
  \gamma_{2*}
  =
  \frac{\epsilon}{3}+O(\epsilon^2),
  \qquad
  \Delta_{\phi^2}
  =
  (2-\epsilon)+\gamma_{2*}
  =
  2-\frac{2}{3}\epsilon+O(\epsilon^2).
\]

\begin{proposition}[Leading \(N=1\) scaling data]
\label{prop:wilson-fisher-leading-scaling-data}
In the one-component scalar theory with interaction normalized as
\(\lambda\phi^4/4!\), the minimal-subtraction pole chart gives
\[
  \gamma_\phi(\lambda)
  =
  \frac{\lambda^2}{12(16\pi^2)^2}+O(\lambda^3),
  \qquad
  \gamma_2(\lambda)
  =
  \frac{\lambda}{16\pi^2}
  +O(\epsilon\lambda,\lambda^2).
\]
On the nonzero small-\(\epsilon\) fixed-point branch
\[
  \lambda_{{\rm MS},*}
  =
  \frac{16\pi^2}{3}\epsilon+O(\epsilon^2),
\]
the elementary field and the even quadratic scalar have dimensions
\[
  \Delta_\phi
  =
  1-\frac{\epsilon}{2}
  +\frac{\epsilon^2}{108}
  +O(\epsilon^3),
  \qquad
  \Delta_{\phi^2}
  =
  2-\frac{2}{3}\epsilon+O(\epsilon^2).
\]
Equivalently,
\[
  \eta=2\gamma_{\phi*}
  =
  \frac{\epsilon^2}{54}+O(\epsilon^3),
  \qquad
  y_t=D-\Delta_{\phi^2}
  =
  2-\frac{\epsilon}{3}+O(\epsilon^2).
\]
\end{proposition}

\begin{proof}
The field statement follows from the pole part of the sunset self-energy
computed above:
\[
  \log Z_\phi^{\rm MS}
  =
  -\frac{\lambda^2}{12(16\pi^2)^2}\frac1\epsilon
  +O(\lambda^3).
\]
Since the engineering part of the beta function is
\(-\epsilon\lambda+O(\lambda^2)\), differentiating
\(\frac12\log Z_\phi^{\rm MS}\) at fixed regulated theory gives
\[
  \gamma_\phi(\lambda)
  =
  \frac12\left(-\epsilon\lambda\right)
  \frac{\partial}{\partial\lambda}
  \left[
    -\frac{\lambda^2}{12(16\pi^2)^2}\frac1\epsilon
  \right]
  +O(\lambda^3)
  =
  \frac{\lambda^2}{12(16\pi^2)^2}+O(\lambda^3).
\]
The quadratic-source statement follows from the nonexceptional insertion
calculation: the factor \(1/2\) in the one-loop graph and the pole of the
massless two-point integral give
\[
  \log Z_{2,\rm src}^{\rm MS}
  =
  -\frac{\lambda}{16\pi^2}\frac1\epsilon
  +O(\lambda\epsilon^0,\lambda^2),
\]
and the same fixed-regulator differentiation gives
\(\gamma_2(\lambda)=\lambda/(16\pi^2)+O(\epsilon\lambda,\lambda^2)\).
The one-loop fixed point is obtained by solving
\[
  -\epsilon\lambda+\frac{3\lambda^2}{16\pi^2}+O(\lambda^3)=0
\]
on the nonzero formal branch.  Substitution into
\[
  \Delta_\phi=\frac{D-2}{2}+\gamma_{\phi*},
  \qquad
  \Delta_{\phi^2}=D-2+\gamma_{2*}
\]
then gives the displayed dimensions.  The formula for \(\eta\) is the
definition of the two-point exponent
\(2\Delta_\phi=D-2+\eta\), and \(y_t=D-\Delta_{\phi^2}\) is
Definition~\ref{def:wilson-fisher-scaling-operator-source-chart} applied to
the source dual to \([\phi^2]\).
\end{proof}

\section{Two-Loop Pole Data and NLO Epsilon Expansion}
\label{sec:wilson-fisher-two-loop-nlo}

The preceding calculation gives the fixed-point branch and the thermal
exponent to leading order.  The next coefficient of \(\nu\), and the next
coefficient of the quartic irrelevant exponent, require the two-loop
minimal-subtraction pole data.  This section stays in the one-component
\(\mathbb Z_2\)-even scalar theory.  The \(O(N)\) generalization is a
different tensor combinatorics problem and is not folded into the notation
here.

Set
\[
  x:=\frac{\lambda_{\rm MS}}{16\pi^2}.
\]
With the action normalized by \(\lambda\phi^4/4!\), the local pole part of the
two-loop four-point calculation is
\begin{equation}
  \frac{g_0}{16\pi^2}
  =
  \mu^\epsilon
  \left[
    x+\frac{3x^2}{\epsilon}
    +x^3\left(\frac{9}{\epsilon^2}
      -\frac{17}{6\epsilon}\right)
    +O(x^4)
  \right].
  \label{eq:wf-two-loop-coupling-pole-map}
\end{equation}
The coefficient \(3\) is the sum of the \(s,t,u\) one-loop bubble channels in
the \(1/4!\) vertex normalization.  The double pole \(9/\epsilon^2\) is not a
new primitive datum: it is fixed by inserting the one-loop local counterterm
into the one-loop four-point graph and by the forest consistency condition
that the cutoff coupling be independent of \(\mu\).  The new two-loop
four-point information is the simple pole
\(-17x^3/(6\epsilon)\).  In a forest subtraction this number is the sum of the
primitive two-loop four-point pole and the local remnants of its subgraph
subtractions, evaluated at any nonexceptional Euclidean subtraction
configuration; locality of the pole part makes it independent of the chosen
configuration.

For the momentum-carrying \(\phi^2\) source used above, let
\(Z_{2,\rm src}^{\rm MS}\) denote the minimal-subtraction source factor
\begin{equation}
  Z_{2,\rm src}^{\rm MS}
  :=
  \frac{\delta\widetilde g_2(p;\mu)}
       {\delta\widetilde g_2(p)} .
  \label{eq:wf-two-loop-phi-two-source-factor}
\end{equation}
Its logarithm has the simple-pole coefficient
\begin{equation}
  \log Z_{2,\rm src}^{\rm MS}
  =
  \frac{-x+\frac{5}{12}x^2}{\epsilon}
  +\hbox{double poles}
  +O(x^3).
  \label{eq:wf-two-loop-phi-two-source-pole}
\end{equation}
The \(-x/\epsilon\) term is the one-loop nonzero-momentum insertion computed
above.  The \(5x^2/(12\epsilon)\) term is the forest-subtracted two-loop
simple pole of the same insertion: the one-loop source counterterm removes
the nested subdivergence, and the remaining local part is proportional to the
same source operator.  The double poles in
\eqref{eq:wf-two-loop-phi-two-source-pole} are determined by lower-order
subdivergences and cancel out of the finite anomalous dimension after the
pole equations are imposed.

For the elementary field, the two-loop sunset computation already gave
\begin{equation}
  \log Z_\phi^{\rm MS}
  =
  -\frac{x^2}{12\epsilon}+O(x^3).
  \label{eq:wf-two-loop-field-log-pole}
\end{equation}
Only this two-loop field pole is needed for \(\eta\) through order
\(\epsilon^2\).  The next coefficient of \(\eta\) would require the three-loop
simple pole in \(Z_\phi^{\rm MS}\), so it is not part of a two-loop
calculation.

\begin{proposition}[Two-loop \(N=1\) Wilson--Fisher epsilon data]
\label{prop:wf-two-loop-nlo-exponents}
In the minimal-subtraction chart with \(x=\lambda_{\rm MS}/(16\pi^2)\), the
two-loop beta function and anomalous dimensions determined by
\eqref{eq:wf-two-loop-coupling-pole-map}--\eqref{eq:wf-two-loop-field-log-pole}
are
\begin{align}
  \beta_x^\epsilon(x)
  &:=
  \frac{\dd x}{\dd\log\mu}
  =
  -\epsilon x+3x^2-\frac{17}{3}x^3+O(x^4),
  \label{eq:wf-two-loop-beta-x}
  \\
  \gamma_2(x)
  &=
  x-\frac56x^2+O(x^3),
  \label{eq:wf-two-loop-gamma-two}
  \\
  \gamma_\phi(x)
  &=
  \frac{x^2}{12}+O(x^3).
  \label{eq:wf-two-loop-gamma-phi-x}
\end{align}
The nonzero small-\(\epsilon\) zero is
\begin{equation}
  x_*
  =
  \frac{\epsilon}{3}
  +\frac{17}{81}\epsilon^2
  +O(\epsilon^3).
  \label{eq:wf-two-loop-fixed-x}
\end{equation}
Consequently
\begin{align}
  \eta
  &=
  \frac{\epsilon^2}{54}+O(\epsilon^3),
  \label{eq:wf-two-loop-eta}
  \\
  \Delta_\phi
  &=
  1-\frac{\epsilon}{2}
  +\frac{\epsilon^2}{108}
  +O(\epsilon^3),
  \label{eq:wf-two-loop-delta-phi}
  \\
  \gamma_{2*}
  &=
  \frac{\epsilon}{3}
  +\frac{19}{162}\epsilon^2
  +O(\epsilon^3),
  \label{eq:wf-two-loop-gamma-two-star}
  \\
  \Delta_{\phi^2}
  &=
  2-\frac{2}{3}\epsilon
  +\frac{19}{162}\epsilon^2
  +O(\epsilon^3),
  \label{eq:wf-two-loop-delta-phi-two}
  \\
  y_t
  &:=
  D-\Delta_{\phi^2}
  =
  2-\frac{\epsilon}{3}
  -\frac{19}{162}\epsilon^2
  +O(\epsilon^3),
  \label{eq:wf-two-loop-y-t}
  \\
  \nu
  &:=
  \frac1{y_t}
  =
  \frac12+\frac{\epsilon}{12}
  +\frac{7}{162}\epsilon^2
  +O(\epsilon^3),
  \label{eq:wf-two-loop-nu}
  \\
  \omega
  &:=
  \left.\frac{\partial\beta_x^\epsilon}{\partial x}\right|_{x=x_*}
  =
  \epsilon-\frac{17}{27}\epsilon^2
  +O(\epsilon^3).
  \label{eq:wf-two-loop-omega}
\end{align}
\end{proposition}

\begin{proof}
First derive the beta function from the pole map, rather than quoting it.  If
\[
  \frac{g_0}{16\pi^2}
  =
  \mu^\epsilon
  \left[
    x+\frac{A x^2}{\epsilon}
    +x^3\left(\frac{A^2}{\epsilon^2}
      +\frac{C}{\epsilon}\right)
    +O(x^4)
  \right],
\]
then differentiating at fixed \(g_0\) gives
\[
  0
  =
  \epsilon F(x,\epsilon)+\beta_x^\epsilon(x)\,\partial_xF(x,\epsilon),
  \qquad
  F(x,\epsilon)
  =
  x+\frac{A x^2}{\epsilon}
  +x^3\left(\frac{A^2}{\epsilon^2}
    +\frac{C}{\epsilon}\right).
\]
Writing
\(\beta_x^\epsilon=-\epsilon x+\beta_1x^2+\beta_2x^3+O(x^4)\) and comparing
powers of \(x\), the \(x^2\) term gives \(\beta_1=A\).  The \(1/\epsilon\)
part of the \(x^3\) equation cancels because the double pole is \(A^2\), and
the finite part gives \(\beta_2=2C\).  In
\eqref{eq:wf-two-loop-coupling-pole-map}, \(A=3\) and \(C=-17/6\), proving
\eqref{eq:wf-two-loop-beta-x}.

For a source factor whose logarithm is
\[
  \log Z(x,\epsilon)=\frac{\ell_1(x)}{\epsilon}+\hbox{higher poles},
\]
the finite anomalous dimension is obtained from
\(\beta_x^\epsilon\partial_x\log Z\).  The pole equations make the higher-pole
contribution cancel the divergent pieces; the finite part is
\[
  -x\,\ell_1'(x)
\]
for a source renormalization and
\[
  -\frac12x\,\ell_1'(x)
\]
for the field factor, because
\(\phi_0=(Z_\phi^{\rm MS})^{1/2}\phi_{\rm MS}\).  Applying this to
\(\ell_1^{(2)}(x)=-x+5x^2/12\) gives
\eqref{eq:wf-two-loop-gamma-two}, while
\(\ell_1^\phi(x)=-x^2/12\) gives
\eqref{eq:wf-two-loop-gamma-phi-x}.

To solve for the fixed point, put
\[
  x_*=a\epsilon+b\epsilon^2+O(\epsilon^3).
\]
Substitution into
\eqref{eq:wf-two-loop-beta-x} gives
\[
  0=(-a+3a^2)\epsilon^2
   +\left(-b+6ab-\frac{17}{3}a^3\right)\epsilon^3
   +O(\epsilon^4).
\]
The nonzero branch has \(a=1/3\), and then the \(\epsilon^3\) equation gives
\(b=17/81\), proving \eqref{eq:wf-two-loop-fixed-x}.

Equations \eqref{eq:wf-two-loop-eta}--\eqref{eq:wf-two-loop-delta-phi-two}
are direct substitutions of \(x_*\) into
\eqref{eq:wf-two-loop-gamma-phi-x} and
\eqref{eq:wf-two-loop-gamma-two}.  The thermal exponent is
\[
  y_t=(4-\epsilon)-\Delta_{\phi^2},
\]
and the expansion of its reciprocal gives
\[
  \frac{1}{2-\epsilon/3-19\epsilon^2/162}
  =
  \frac12+\frac{\epsilon}{12}+\frac{7\epsilon^2}{162}
  +O(\epsilon^3).
\]
Finally,
\[
  \partial_x\beta_x^\epsilon
  =
  -\epsilon+6x-17x^2+O(x^3),
\]
and substituting \(x_*\) gives
\[
  -\epsilon+6\left(\frac{\epsilon}{3}
    +\frac{17}{81}\epsilon^2\right)
  -17\left(\frac{\epsilon}{3}\right)^2
  =
  \epsilon-\frac{17}{27}\epsilon^2+O(\epsilon^3).
\]
\end{proof}

\section{Critical Surface and the Mass Exponent}

The mass perturbation is the deformation generated by the renormalized
operator \([O_2]_\mu\).  Let \(t\) be a renormalized coefficient for this
deformation, so that near the fixed point
\[
  \delta S_t
  =
  \int \dd^D x\,t\, [O_2]_\mu(x).
\]
The dimension of \(t\) at the fixed point is
\[
  y_t
  =
  D-\Delta_{\phi^2}.
\]
Using the result above,
\[
  y_t
  =
  (4-\epsilon)
  -
  \left(
    2-\frac{2}{3}\epsilon
    +\frac{19}{162}\epsilon^2
  \right)
  +
  O(\epsilon^3)
  =
  2-\frac{\epsilon}{3}
  -\frac{19}{162}\epsilon^2
  +O(\epsilon^3).
\]
This identification uses the fact that the thermal source is already a
linearized scaling coordinate in the \(\mathbb Z_2\)-even scalar sector at the
order being computed.  More explicitly, let \(g^A\) denote source coordinates
for a finite list of even scalar local operators \(O_A\), with the identity
operator and equation-of-motion or total-derivative redundant directions
removed.  The linearized source RG at the fixed point has the form
\[
  \frac{\dd g^A}{\dd\log\mu}
  =
  -M^A{}_B\,g^B+O(g^2),
\]
and the relevant exponents are the eigenvalues of \(M^A{}_B\), after
diagonalizing possible operator mixing.  In the one-component scalar theory at
this order, the even scalar sector containing the mass deformation has a single
relevant nontrivial basis element, represented by \([O_2]_\mu=[\phi^2]_\mu\).
The identity source is the vacuum-energy coordinate and \(\phi^4\) is the
quartic direction, whose fixed-point eigenvalue is the positive irrelevant
exponent \(\omega=\epsilon-\frac{17}{27}\epsilon^2+O(\epsilon^3)\).  Hence the
thermal block is one-dimensional and its eigenvalue is \(D-\Delta_{\phi^2}\).
In a theory with another even scalar of the same
quantum numbers and degenerate dimension, \(t\) would have to be replaced by
the corresponding eigen-coordinate of \(M^A{}_B\), and the formula above would
refer to that eigenvalue rather than to a chosen basis operator.
The corresponding dimensionless coordinate
\[
  \tau(\mu)=\mu^{-y_t}t
\]
obeys, to linear order at the fixed point,
\[
  \frac{\dd\tau}{\dd\log\mu}
  =
  -y_t\,\tau+O(\tau^2,\tau(\lambda-\lambda_*)).
\]
Thus \(\tau\) grows when \(\mu\) is lowered.  In the \(\mathbb Z_2\)-even
finite chart and to linear order at the fixed point, the condition
\(\tau=0\) is the codimension-one relevant-coordinate condition, with the
quartic coordinate flowing to \(\lambda_*\) along it.  The corresponding
nonlinear statement that the tuned set is a \(C^k\) codimension-one
submanifold is the finite-reference implicit-function construction of
Chapter~\ref{ch:vol2-wilsonian-effective-actions} applied to the
\(\mathbb Z_2\)-even RG chart, after the endpoint map and its nondegeneracy
have been established.  Choosing \(m_0^2\) so that the long-distance theory is
massless is the microscopic way of imposing this endpoint condition.

If the \(\mathbb Z_2\) symmetry is not imposed, an odd deformation
\[
  \delta S_h
  =
  \int \dd^D x\,h\,[\phi]_\mu(x)
\]
is present.  Its fixed-point eigenvalue is
\[
  y_h
  =
  D-\Delta_\phi
  =
  3-\frac{\epsilon}{2}
  -\frac{\epsilon^2}{108}
  +O(\epsilon^3),
\]
so \(h\) is also relevant.  The statement that the massless critical surface
has one relevant tuning condition is therefore a statement about the
\(\mathbb Z_2\)-even subspace; in the local chart including \(h\), both
\(h\) and \(t\) must be set to zero to reach the symmetric Wilson-Fisher
fixed point.  A codimension statement again uses the Banach-chart hypotheses
of the finite-reference implicit-function construction.

Let a trajectory start at a reference scale \(\mu_0\) with small nonzero
\(\tau_0=\tau(\mu_0)\).  Linearized running gives
\[
  \tau(\mu)
  =
  \tau_0\left(\frac{\mu}{\mu_0}\right)^{-y_t}.
\]
The scale at which \(|\tau(\mu)|\) becomes order one sets the inverse
correlation length:
\[
  \xi^{-1}
  \propto
  \mu_0\,|\tau_0|^{1/y_t}.
\]
Therefore the correlation-length exponent in this perturbative calculation is
\[
  \nu=\frac1{y_t}
  =
  \frac12+\frac{\epsilon}{12}
  +\frac{7}{162}\epsilon^2
  +O(\epsilon^3).
\]
The proportionality constant in \(\xi^{-1}\) depends on the normalization of
the thermal scaling field \(\tau\); the exponent does not.

Figure~\ref{fig:volume-ii-wilson-fisher-critical-surface} records the
\(\mathbb Z_2\)-even critical surface in the local coordinates used here: the
thermal direction is relevant, while the critical trajectory flows into the
interacting fixed point.

\begin{figure}[H]
  \centering
  \resizebox{0.78\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \node[font=\small] at (0,2.55) {\(\mathbb Z_2\)-even slice \(h=0\)};
    \draw[-{Latex[length=2mm]}, thick] (-3.2,0)--(3.4,0)
      node[below] {\(\lambda\)};
    \draw[-{Latex[length=2mm]}, thick] (0,-2.1)--(0,2.3)
      node[left] {\(\tau\)};
    \draw[purple!80!black, very thick] (-2.8,0)--(3.0,0);
    \node[purple!80!black, above] at (-2.15,0.28)
      {critical surface};
    \fill[black] (-2.15,0) circle (2pt);
    \node[below] at (-2.15,-0.22) {Gaussian};
    \fill[blue!75!black] (1.15,0) circle (2.6pt);
    \node[blue!75!black, below] at (1.15,-0.2) {\(\lambda_*\)};
    \foreach \x in {-1.35,-0.35,0.55} {
      \draw[-{Latex[length=2mm]}, purple!80!black, very thick]
        (\x,0.25)--++(0.48,0);
    }
    \draw[-{Latex[length=2mm]}, red!75!black, very thick,
      dash pattern=on 4pt off 2pt]
      (1.15,0.12)--(1.15,1.05);
    \draw[-{Latex[length=2mm]}, red!75!black, very thick,
      dash pattern=on 4pt off 2pt]
      (1.15,-0.12)--(1.15,-1.05);
    \node[red!75!black, align=center] at (2.35,1.15)
      {relevant\\mass direction};
    \node[blue!70!black, align=center] at (-1.75,-1.25)
      {\(y_t=D-\Delta_{\phi^2}\)\\
       \(\nu=1/y_t\)};
  \end{tikzpicture}
  }
  \caption{In the \(\mathbb Z_2\)-even finite chart, the massless endpoint
  condition removes the relevant thermal direction.  The dashed transverse
  arrows show the mass perturbation, the direction dual to \([O_2]\), with
  eigenvalue
  \(y_t=D-\Delta_{\phi^2}\).}
  \label{fig:volume-ii-wilson-fisher-critical-surface}
\end{figure}

\section{Correlators and Local Composite Fields}

At the fixed point, the running of the renormalized momentum-space operator is
\[
  [\widetilde O_2(p)]_{\mu=|p|}
  =
  \left(\frac{|p|}{\mu_0}\right)^{-\gamma_{2*}}
  \widetilde O_2(p),
\]
where \(\mu_0\) is a fixed reference scale.  The correlator of renormalized
operators is finite and, by dimensional analysis,
\[
  \left\langle
    [\widetilde O_2(p)]_{\mu=|p|}
    [\widetilde O_2(q)]_{\mu=|q|}
  \right\rangle
  \propto
  \delta^{(4-\epsilon)}(p+q)\,|p|^{-\epsilon}.
\]
Consequently
\[
  \mu_0^{2\gamma_{2*}}
  \left\langle
    \widetilde O_2(p)\widetilde O_2(q)
  \right\rangle
  \propto
  \delta^{(4-\epsilon)}(p+q)\,|p|^{-\epsilon+2\gamma_{2*}}.
\]
Fourier transformation gives the position-space scaling law
\[
  \mu_0^{2\gamma_{2*}}
  \left\langle
    O_2(x)O_2(0)
  \right\rangle
  \propto
  \frac{1}{|x|^{2\Delta_{\phi^2}}},
  \qquad
  \Delta_{\phi^2}=2-\epsilon+\gamma_{2*}.
\]

Figure~\ref{fig:volume-ii-phi-two-correlator-scaling} displays the invariant
content of this calculation.  Once a common normalization is chosen at
\(\mu_0|x|=1\), the scaling dimension is the slope of the correlator on
logarithmic axes.  The anomalous dimension \(\gamma_{2*}\) changes that slope
by \(2\gamma_{2*}\), not by introducing a new scale.

\begin{figure}[H]
  \centering
  \resizebox{0.94\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      axis/.style={-{Latex[length=2mm]},line width=0.65pt},
      curve/.style={line width=1.1pt},
      note/.style={align=center,inner sep=2pt}
    }

    \node[font=\small] at (3.9,4.95)
      {fixed-point position-space correlator on logarithmic axes};
    \draw[axis] (-0.10,0)--(7.85,0)
      node[below] {\(X=\log(\mu_0|x|)\)};
    \draw[axis] (0,-0.10)--(0,4.55)
      node[left] {\(Y=-\log(G_2(|x|)/A)\)};
    \node[anchor=east] at (-0.06,0) {\(0\)};
    \node[below] at (0,-0.08) {\(\mu_0|x|=1\)};

    \draw[curve,gray!65!black]
      (0,0)--(6.7,2.95);
    \draw[curve,blue!70!black]
      (0,0)--(6.7,3.72);
    \draw[dashed,gray!50] (5.75,0)--(5.75,3.42);
    \draw[{Latex[length=1.8mm]}-{Latex[length=1.8mm]},red!70!black,
      line width=0.75pt] (5.75,2.53)--(5.75,3.20);
    \node[red!70!black,anchor=west] at (5.90,2.86)
      {\(2\gamma_{2*}X\)};

    \node[gray!65!black,anchor=west,fill=white,inner sep=1.5pt]
      at (3.70,2.24)
      {\(\displaystyle
        Y_{\rm G}(X)=2(2-\epsilon)\,X\)};
    \node[blue!70!black,anchor=west,fill=white,inner sep=1.5pt]
      at (3.30,3.83)
      {\(\displaystyle
        Y_{\rm WF}(X)=2(2-\epsilon+\gamma_{2*})\,X\)};

    \node[note] at (2.00,3.88)
      {\(\displaystyle
        G_2(r)=A(\mu_0 r)^{-2\Delta_{\phi^2}}\),\\
       \(\displaystyle
        \Delta_{\phi^2}=2-\epsilon+\gamma_{2*}\)};
    \node[note,blue!70!black] at (6.20,0.72)
      {long-distance decay is faster\\
       by the slope shift \(2\gamma_{2*}\)};
  \end{tikzpicture}
  }
  \caption{Logarithmic plot of the \(\phi^2\) two-point function at a fixed
  point, with \(G_2(r)=\mu_0^{2\gamma_{2*}}\langle O_2(x)O_2(0)\rangle\),
  \(r=|x|\), and a common normalization \(G_2(1/\mu_0)=A\).  The Gaussian
  scalar composite has slope \(2(2-\epsilon)\).  The Wilson--Fisher scaling
  operator has slope \(2\Delta_{\phi^2}=2(2-\epsilon+\gamma_{2*})\).  Thus the
  anomalous dimension appears as a slope shift \(2\gamma_{2*}\) in the
  log-log correlator.}
  \label{fig:volume-ii-phi-two-correlator-scaling}
\end{figure}

There is a related locality point.  The object obtained by choosing the
subtraction scale separately for each Fourier mode,
\[
  [O_2(x)]_{\rm mode}
  =
  \int\frac{\dd^{4-\epsilon}p}{(2\pi)^{4-\epsilon}}\,
  e^{ip\cdot x}\,
  [\widetilde O_2(p)]_{\mu=|p|},
\]
is not a local field.  By the running formula it equals
\[
  \int\frac{\dd^{4-\epsilon}p}{(2\pi)^{4-\epsilon}}\,
  e^{ip\cdot x}
  \left(\frac{|p|}{\mu_0}\right)^{-\gamma_{2*}}
  \widetilde O_2(p),
\]
and multiplication by the non-polynomial function
\(|p|^{-\gamma_{2*}}\) is nonlocal in position space.  The local scaling
field is obtained by fixing the reference normalization once and for all.

The notation \(O_2=\phi^2\) is a label for the renormalized local field in this
operator family; it is not a definition by an unregulated product at
coincident points.  A point-splitting representative has the form
\[
  \mu_0^{\gamma_{2*}}O_2(x)
  \propto
  \lim_{x'\to x}
  |x'-x|^{2\gamma_{\phi*}-\gamma_{2*}}
  \left[
    \hat\phi(x')\hat\phi(x)
    -
    \left\langle\hat\phi(x')\hat\phi(x)\right\rangle
  \right],
\]
where the subtraction removes the singular two-point function before the
coincident-point limit is taken.

\section{Descendants and Irrelevant Directions}

The cubic operator is not independent at the interacting fixed point.  In the
renormalized operator algebra, the equation of motion gives a descendant
relation with a coupling-dependent normalization.  If the renormalized
dimensionful quartic coupling is written as
\[
  g_R(\mu)=\mu^\epsilon\lambda(\mu),
\]
then, away from contact terms and up to the overall sign convention for
\(\Box\), the renormalized equation of motion gives
\[
  \Box\phi(x)=\frac{g_R(\mu)}{3!}\,[\phi^3(x)]_\mu .
\]
At the fixed point this is equivalently
\[
  [\phi^3(x)]_\mu
  =
  \frac{3!}{\mu^\epsilon\lambda_*}\,\Box\phi(x),
\]
again modulo contact terms and finite normalizations of the chosen composite
operator.  If the proportionality coefficient is denoted by
\(\kappa_*(\mu)\),
\[
  [\phi^3(x)]_\mu=\kappa_*(\mu)\,\Box\phi(x),
\]
then the action normalization above gives, at the classical level,
\[
  \kappa_{\rm cl}(\mu)
  =
  \frac{3!}{g_R(\mu)}
  =
  \frac{6}{g_R(\mu)}.
\]
A perturbative finite normalization of the composite operator replaces this
by
\[
  \kappa_*(\mu)
  =
  \frac{6}{g_R(\mu)}
  \bigl(1+O(g_R(\mu))\bigr)
  =
  \frac{6}{g_R(\mu)}+O(1),
\]
with the \(O(1)\) term depending on the composite-operator convention.  The
statement invariant under such choices is that \(O_3\) is an
equation-of-motion descendant of \(O_1=\phi\).  Its dimension is
\[
  \Delta_{\phi^3}
  =
  \Delta_\phi+2.
\]

The quartic operator \(O_4=\phi^4\) is instead the operator associated with
motion in the quartic-coupling direction.  As for \(O_2\), one may first use a
momentum-dependent source,
\[
  \delta S_4
  =
  \int\dd^D x\,\delta g_4(x)\,\phi^4(x)
  =
  \int\frac{\dd^Dp}{(2\pi)^D}\,
  \delta\widetilde g_4(p)\,\widetilde O_4(-p).
\]
The renormalized source is extracted from a four-point 1PI insertion with
external momenta of order \(\mu\) and source momentum \(p=O(\mu)\).  In this
case one may set \(p=0\): the massless infrared singularity that obstructed
the zero-momentum \(O_2\) subtraction is absent for this insertion.  The
momentum-dependent source definition therefore agrees with the earlier
\(x\)-independent coupling-deformation definition.
A small source perturbation obeys the linearized RG equation
\[
  \frac{\dd\,\delta g_4(\mu)}{\dd\log\mu}
  =
  \frac{\partial\beta^\epsilon(\lambda(\mu))}
       {\partial\lambda(\mu)}
  \,\delta g_4(\mu).
\]
Therefore
\[
  \omega
  =
  \left.
  \frac{\partial\beta^\epsilon(\lambda)}{\partial\lambda}
  \right|_{\lambda=\lambda_*}
  =
  \left.
  \left(
    -\epsilon+\frac{3\lambda}{8\pi^2}
    -\frac{17\lambda^2}{(16\pi^2)^2}
    +O(\lambda^3)
  \right)
  \right|_{\lambda=16\pi^2
    \left(\frac{\epsilon}{3}+\frac{17}{81}\epsilon^2\right)
    +O(\epsilon^3)}
  =
  \epsilon-\frac{17}{27}\epsilon^2+O(\epsilon^3),
\]
which is positive.  This positivity is the same statement as infrared
attractiveness of the quartic direction when the flow is followed toward
smaller \(\mu\).
The scaling dimension is therefore
\[
  \Delta_{\phi^4}
  =
  D+\omega
  =
  4-\frac{17}{27}\epsilon^2+O(\epsilon^3)
  >
  D.
\]
Thus \(\phi^4\) is irrelevant at the interacting fixed point.  This statement
is about perturbations around the fixed point; it is compatible with the fact
that the quartic interaction was needed to produce that fixed point.

Figure~\ref{fig:volume-ii-wilson-fisher-operator-spectrum} summarizes the local
operator-spectrum conclusion of this perturbative calculation: descendants,
relevant primaries, and the quartic perturbation's irrelevant fixed-point
eigenvalue occupy distinct roles.

\begin{figure}[H]
  \centering
  \resizebox{0.92\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \begin{scope}
      \node[font=\small, align=center] at (0,2.35)
        {\(\phi^3\) as descendant};
      \node[draw=black, rounded corners, inner sep=6pt] at (-1.05,0.7) {\(\phi^3\)};
      \node[draw=black, rounded corners, inner sep=6pt] at (1.2,0.7) {\(\Box\phi\)};
      \draw[-{Latex[length=2mm]}, thick] (-0.45,0.7)--(0.55,0.7);
      \node[above] at (0.05,0.92) {EOM};
      \node[blue!70!black, align=center] at (0,-0.55)
        {\(\Delta_{\phi^3}=\Delta_\phi+2\)};
    \end{scope}

    \begin{scope}[shift={(5.9,0)}]
      \node[font=\small, align=center] at (0,2.35)
        {\(\phi^4\) as irrelevant direction};
      \fill[gray!18, draw=black, thick] (0,0.7) circle (0.5);
      \node at (0,0.7) {1PI};
      \foreach \a in {35,145,215,325} {
        \draw[thick] ({0.5*cos(\a)},{0.7+0.5*sin(\a)})
          --++({0.92*cos(\a)},{0.92*sin(\a)});
      }
      \fill[red!75!black] (0,1.58) circle (2.5pt);
      \draw[red!75!black, thick] (0,1.58)--(0,1.2);
      \node[red!75!black, above] at (0,1.84) {\(\delta g_4\)};
      \node[blue!70!black, align=center] at (0,-0.55)
        {\(\omega=(\beta^\epsilon)'(\lambda_*)=\epsilon-\frac{17}{27}\epsilon^2+\cdots\)\\
         \(\Delta_{\phi^4}>D\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The cubic operator is an equation-of-motion descendant.  The quartic
  operator is independent, but its dimension is larger than \(D\) at the
  interacting fixed point, so changing its coefficient is an irrelevant
  deformation.}
  \label{fig:volume-ii-wilson-fisher-operator-spectrum}
\end{figure}

The low-lying scalar dimensions are summarized as follows.  In the third
column, \(O_{\rm irr}\) denotes the leading nonredundant
\(\mathbb Z_2\)-even scalar irrelevant deformation after the identity and the
energy operator have been removed.  In the \(4-\epsilon\) perturbative chart
this direction is represented by \(\phi^4\); in lower-dimensional fixed-point
CFTs the notation \(O_{\rm irr}\) refers to the corresponding scaling operator,
not to an unrenormalized fourth power of the spin field.
\[
  \begin{array}{c|ccc}
    & \Delta_\phi & \Delta_{\phi^2} & \Delta_{O_{\rm irr}} \\
    \hline
    d=4-\epsilon
      & 1-\frac{\epsilon}{2}
        +\frac{\epsilon^2}{108}+O(\epsilon^3)
      & 2-\frac{2}{3}\epsilon
        +\frac{19}{162}\epsilon^2+O(\epsilon^3)
      & 4-\frac{17}{27}\epsilon^2+O(\epsilon^3) \\
    d=3
      & 0.518148806(24)
      & 1.41262528(29)
      & 3.82951(61) \\
    d=2
      & \frac18
      & 1
      & 4
  \end{array}
\]
The first row is the perturbative epsilon expansion.  The \(d=3\) and \(d=2\)
rows are data about the corresponding fixed points themselves; they are not
deduced by substituting \(\epsilon=1\) or \(\epsilon=2\) into the finite
truncation displayed in the first row.  In the \(d=3\) row the fixed-point
fields are conventionally denoted
\(\sigma,\varepsilon,\varepsilon'\): \(\phi,\phi^2,\phi^4\) are the Landau
representatives of the corresponding scaling directions, not definitions by
unrenormalized powers at the fixed point.  The displayed numerical values are
conformal-bootstrap data for the three-dimensional Ising CFT.  The quoted
\(\Delta_\sigma\) and \(\Delta_\varepsilon\) are the mixed-correlator values
reported in Chang, Dommes, Erramilli, Homrich, Kravchuk, Liu, Mitchell,
Poland, and Simmons-Duffin, \emph{Bootstrapping the 3d Ising Stress Tensor},
arXiv:2411.15300; earlier precision islands for these two dimensions were
obtained by El-Showk, Paulos, Poland, Rychkov, Simmons-Duffin, and Vichi,
\emph{Phys. Rev. D} \textbf{86} (2012) 025022, arXiv:1203.6064;
\emph{J. Stat. Phys.} \textbf{157} (2014) 869--914, arXiv:1403.4545; and by
Kos, Poland, Simmons-Duffin, and Vichi, \emph{JHEP} \textbf{08} (2016) 036,
arXiv:1603.04436.  The value in the third column is the dimension of the
leading \(\mathbb Z_2\)-even scalar irrelevant primary \(\varepsilon'\), whose
Landau representative is the quartic direction; the number displayed here is
the spectral estimate quoted in Simmons-Duffin, \emph{JHEP} \textbf{03}
  (2017) 086, arXiv:1612.08471.  The mixed-correlator crossing and positivity
system underlying the \(\Delta_\sigma,\Delta_\varepsilon\) island is derived
from the higher-dimensional conformal-block conventions in
Section~\ref{sec:ising-mixed-correlator-bootstrap-system}; the numerical
island itself remains external high-precision input rather than a theorem
proved in this perturbative Wilson--Fisher chapter.

The \(d=2\) row is exact CFT data.  The critical Ising CFT is the unitary
minimal model \(\mathcal M(3,4)\) with Virasoro primaries
\(\sigma\) and \(\varepsilon\) of weights
\[
  (h_\sigma,\bar h_\sigma)=\left(\frac1{16},\frac1{16}\right),
  \qquad
  (h_\varepsilon,\bar h_\varepsilon)=\left(\frac12,\frac12\right),
\]
so \(\Delta_\sigma=h_\sigma+\bar h_\sigma=1/8\) and
\(\Delta_\varepsilon=1\).  The scalar irrelevant entry of dimension \(4\) is
the \(T\bar T\) Virasoro descendant of the identity, the two-dimensional
continuation of the leading even irrelevant lattice-scalar direction recorded in
the third column.  It is not the dimension of an unrenormalized \(\sigma^4\)
operator.  These exact statements are the conformal-field-theoretic form of the
Onsager critical Ising solution, and follow from the minimal-model
classification of Belavin, Polyakov, and Zamolodchikov.

\section{\texorpdfstring{The \(O(N)\) Wilson--Fisher Family}{The O(N) Wilson-Fisher Family}}
\label{sec:on-wilson-fisher-family}

\begin{definition}[\(O(N)\) quartic scalar chart]
\label{def:on-quartic-scalar-chart}
For a positive integer \(N\), let
\(\phi=(\phi^1,\ldots,\phi^N)\) be an \(O(N)\)-vector scalar.  Indices are
raised and lowered with \(\delta_{ij}\), repeated \(O(N)\) indices are
summed, and
\[
  \phi^2:=\delta_{ij}\phi^i\phi^j .
\]
The \(O(N)\)-invariant quartic theory in \(D=4-\epsilon\) is
\begin{equation}
  S_E[\phi]
  =
  \int \dd^D x\,
  \left[
    \frac12\partial_\mu\phi^i\partial_\mu\phi^i
    +\frac12 r_0\,\phi^i\phi^i
    +\frac{g_0}{4!}\,(\phi^i\phi^i)^2
  \right].
  \label{eq:on-vector-phi-four-action}
\end{equation}
The quartic vertex tensor associated with the normalization in
\eqref{eq:on-vector-phi-four-action} is
\begin{equation}
  V_{ij;kl}
  =
  \frac{g_0}{3}
  \left(
    \delta_{ij}\delta_{kl}
    +\delta_{ik}\delta_{jl}
    +\delta_{il}\delta_{jk}
  \right).
  \label{eq:on-quartic-vertex-tensor}
\end{equation}
For \(N=1\) the tensor in parentheses equals \(3\), so this is the same
\(g_0\phi^4/4!\) normalization used above.  The singlet quadratic source is
coupled to
\[
  S(x):=\frac12\phi^i(x)\phi^i(x).
\]
The factor \(1/2\) is a source convention; it does not affect the scaling
dimension.

Set again
\[
  x:=\frac{\lambda_{\rm MS}}{16\pi^2}.
\]
\end{definition}

The one-component theory is the \(N=1\) member of this family.  In the
minimal-subtraction chart determined by the singlet quartic vertex
\eqref{eq:on-quartic-vertex-tensor}, the two-loop local pole map is
\begin{equation}
  \frac{g_0}{16\pi^2}
  =
  \mu^\epsilon
  \left[
    x
    +\frac{N+8}{3}\frac{x^2}{\epsilon}
    +x^3\left\{
      \frac{(N+8)^2}{9\epsilon^2}
      -\frac{3N+14}{6\epsilon}
    \right\}
    +O(x^4)
  \right].
  \label{eq:on-two-loop-coupling-pole-map}
\end{equation}
The coefficient \(N+8\) has a concrete origin.  In a one-loop four-point
graph, the closed internal index loop gives the \(N\) term, while the exchange
contractions among the three tensor structures in
\eqref{eq:on-quartic-vertex-tensor} give the \(8\) term.  The double pole in
\eqref{eq:on-two-loop-coupling-pole-map} is fixed by this one-loop tensor
contraction and the forest equations.  The new two-loop tensor datum is the
simple-pole coefficient \(-(3N+14)/6\).

The singlet source and field pole factors are
\begin{align}
  \log Z_{S,\rm src}^{\rm MS}
  &=
  \frac{
    -\frac{N+2}{3}x+\frac{5(N+2)}{18}x^2
  }{\epsilon}
  +\hbox{double poles}
  +O(x^3),
  \label{eq:on-source-pole}
  \\
  \log Z_\phi^{\rm MS}
  &=
  -\frac{N+2}{36}\frac{x^2}{\epsilon}
  +O(x^3).
  \label{eq:on-field-pole}
\end{align}
The factor \(N+2\) is the index contraction of a singlet \(\phi^2\) insertion
with the quartic tensor.  The same pole-equation argument used in the
one-component calculation gives the finite anomalous dimensions.

\begin{proposition}[\(O(N)\) Wilson--Fisher data through two loops]
\label{prop:on-wilson-fisher-two-loop-data}
In the MS coordinate \(x=\lambda_{\rm MS}/(16\pi^2)\), the singlet
\(O(N)\)-invariant quartic chart has
\begin{align}
  \beta_x^\epsilon
  &=
  -\epsilon x
  +\frac{N+8}{3}x^2
  -\frac{3N+14}{3}x^3
  +O(x^4),
  \label{eq:on-beta-x}
  \\
  \gamma_S(x)
  &=
  \frac{N+2}{3}x
  -\frac{5(N+2)}{18}x^2
  +O(x^3),
  \label{eq:on-gamma-S}
  \\
  \gamma_\phi(x)
  &=
  \frac{N+2}{36}x^2
  +O(x^3).
  \label{eq:on-gamma-phi}
\end{align}
The nonzero perturbative fixed-point coordinate is
\begin{equation}
  x_*
  =
  \frac{3}{N+8}\epsilon
  +\frac{9(3N+14)}{(N+8)^3}\epsilon^2
  +O(\epsilon^3).
  \label{eq:on-x-star}
\end{equation}
The singlet thermal exponent, field anomalous dimension, and leading quartic
irrelevant exponent are
\begin{align}
  \eta
  &=
  \frac{N+2}{2(N+8)^2}\epsilon^2
  +O(\epsilon^3),
  \label{eq:on-eta}
  \\
  \Delta_\phi
  &=
  1-\frac{\epsilon}{2}
  +\frac{N+2}{4(N+8)^2}\epsilon^2
  +O(\epsilon^3),
  \label{eq:on-delta-phi}
  \\
  \Delta_S
  &=
  2-\frac{6}{N+8}\epsilon
  +\frac{(N+2)(13N+44)}{2(N+8)^3}\epsilon^2
  +O(\epsilon^3),
  \label{eq:on-delta-S}
  \\
  y_t
  &:=
  D-\Delta_S
  =
  2-\frac{N+2}{N+8}\epsilon
  -\frac{(N+2)(13N+44)}{2(N+8)^3}\epsilon^2
  +O(\epsilon^3),
  \label{eq:on-y-t}
  \\
  \nu
  &:=
  \frac1{y_t}
  =
  \frac12
  +\frac{N+2}{4(N+8)}\epsilon
  +\frac{(N+2)(N^2+23N+60)}{8(N+8)^3}\epsilon^2
  +O(\epsilon^3),
  \label{eq:on-nu}
  \\
  \omega
  &:=
  \left.\partial_x\beta_x^\epsilon\right|_{x=x_*}
  =
  \epsilon
  -\frac{3(3N+14)}{(N+8)^2}\epsilon^2
  +O(\epsilon^3).
  \label{eq:on-omega}
\end{align}
\end{proposition}

\begin{proof}
Let
\[
  A=\frac{N+8}{3},
  \qquad
  C=-\frac{3N+14}{6}.
\]
Equation \eqref{eq:on-two-loop-coupling-pole-map} has the same algebraic form
as the one-component pole map:
\[
  \frac{g_0}{16\pi^2}
  =
  \mu^\epsilon
  \left[
    x+\frac{A x^2}{\epsilon}
    +x^3\left(\frac{A^2}{\epsilon^2}
      +\frac{C}{\epsilon}\right)
    +O(x^4)
  \right].
\]
Differentiation at fixed \(g_0\) gives
\(\beta_x^\epsilon=-\epsilon x+Ax^2+2Cx^3+O(x^4)\), which is
\eqref{eq:on-beta-x}.  The simple-pole coefficient in
\eqref{eq:on-source-pole} is
\[
  \ell_1^S(x)
  =
  -\frac{N+2}{3}x
  +\frac{5(N+2)}{18}x^2 .
\]
The finite source anomalous dimension is \(-x(\ell_1^S)'(x)\), up to
terms beyond two loops, giving \eqref{eq:on-gamma-S}.  Similarly the field
simple-pole coefficient
\(\ell_1^\phi(x)=-(N+2)x^2/36\) gives
\(-\frac12x(\ell_1^\phi)'(x)\), namely \eqref{eq:on-gamma-phi}.

Solving \(\beta_x^\epsilon(x_*)=0\) with
\(x_*=a\epsilon+b\epsilon^2+O(\epsilon^3)\) gives
\[
  a=\frac1A=\frac{3}{N+8},
  \qquad
  b=-\frac{2C}{A^3}
  =
  \frac{9(3N+14)}{(N+8)^3},
\]
which proves \eqref{eq:on-x-star}.  Substituting
\eqref{eq:on-x-star} into \eqref{eq:on-gamma-phi} gives
\(\eta=2\gamma_\phi(x_*)\), hence \eqref{eq:on-eta} and
\eqref{eq:on-delta-phi}.  Substitution into \eqref{eq:on-gamma-S} gives
\[
  \gamma_S(x_*)
  =
  \frac{N+2}{N+8}\epsilon
  +\frac{(N+2)(13N+44)}{2(N+8)^3}\epsilon^2
  +O(\epsilon^3).
\]
Since the singlet \(S\) has engineering dimension \(D-2=2-\epsilon\),
\(\Delta_S=2-\epsilon+\gamma_S(x_*)\).  Equations
\eqref{eq:on-delta-S}--\eqref{eq:on-nu} follow by expanding
\(y_t=D-\Delta_S=2-\gamma_S(x_*)\) and then its reciprocal.  Finally,
\[
  \partial_x\beta_x^\epsilon
  =
  -\epsilon
  +\frac{2(N+8)}{3}x
  -(3N+14)x^2
  +O(x^3),
\]
and substituting \eqref{eq:on-x-star} gives \eqref{eq:on-omega}.
\end{proof}

For \(N=1\), Proposition \ref{prop:on-wilson-fisher-two-loop-data} reduces
to Proposition \ref{prop:wf-two-loop-nlo-exponents}.  For \(N>1\) there are
additional non-singlet operator sectors.  The symmetric traceless tensor
\[
  T^{ij}:=\phi^i\phi^j-\frac{\delta^{ij}}{N}\phi^k\phi^k
\]
and the vector \(\phi^i\) have their own source-RG blocks.  The formulas above
concern the singlet thermal direction and the elementary vector field.  A
complete operator-spectrum analysis must diagonalize each \(O(N)\)
representation sector separately.

\subsection{\texorpdfstring{Large \(N\), the Spherical Model, and the
Sigma-Model Chart}{Large N, the Spherical Model, and the Sigma-Model
Chart}}
\label{subsec:on-large-n-spherical-sigma}

The limit \(N\to\infty\) is not obtained by holding \(x\) fixed.  The fixed
point in \eqref{eq:on-x-star} has \(x_*=O(1/N)\), while the collective
coordinate
\[
  u:=Nx
\]
has a finite limit.  Multiplying \eqref{eq:on-beta-x} by \(N\) and putting
\(x=u/N\) gives
\begin{equation}
  \beta_u^\epsilon
  =
  -\epsilon u+\frac13u^2
  +O\left(\frac1N\right),
  \qquad
  u_*=3\epsilon+O\left(\frac1N\right).
  \label{eq:on-large-n-u-beta}
\end{equation}
Thus the quartic coordinate of each finite-\(N\) fixed point tends to zero,
but the singlet collective interaction stays finite.

The nonperturbative large-\(N\) limit is best derived before expanding around
four dimensions.  Work with a finite ultraviolet cutoff and an \(O(N)\)-invariant
potential of the form
\begin{equation}
  S[\phi]
  =
  \int \dd^D x\,
  \left[
    \frac12\partial_\mu\phi^i\partial_\mu\phi^i
    +N\,U\left(\rho\right)
  \right],
  \qquad
  \rho:=\frac{\phi^i\phi^i}{N}.
  \label{eq:on-large-n-potential-action}
\end{equation}
At finite cutoff, introduce fields \(\rho(x)\) and \(\sigma(x)\) by a
Fourier representation of the constraint
\(\rho(x)=\phi^i(x)\phi^i(x)/N\).  The contour of \(\sigma\) is chosen so
that the finite-dimensional regulated integral is convergent and may be
deformed to the steepest-descent contour.  After integrating over the
\(N\) vector components, the effective action per component is
\begin{equation}
  \frac1N S_{\rm eff}[\rho,\sigma]
  =
  \int \dd^D x\,
  \left[
    U(\rho)-\frac12\sigma\rho
  \right]
  +\frac12\operatorname{Tr}_\Lambda
    \log(-\partial^2+\sigma).
  \label{eq:on-large-n-effective-action}
\end{equation}
For constant saddle fields, write \(\sigma=m^2\).  The saddle equations are
\begin{equation}
  U'(\rho)=\frac12m^2,
  \qquad
  \rho=\Omega_D(m),
  \qquad
  \Omega_D(m)
  :=
  \int^\Lambda\frac{\dd^D p}{(2\pi)^D}\frac1{p^2+m^2},
  \label{eq:on-large-n-gap-equations}
\end{equation}
in the symmetric phase.  If an ordered component
\(\langle\phi^1\rangle=M\) is allowed, the second equation becomes
\[
  \rho=\frac{M^2}{N}+\Omega_D(m),
  \qquad
  m^2M=0.
\]
The critical point is obtained by tuning the relevant parameter in \(U\) so
that \(m=0\) is a saddle:
\[
  \rho_c=\Omega_D(0),
  \qquad
  U'(\rho_c)=0.
\]
For \(2<D<4\), the small-\(m\) expansion of the regulated integral has a
universal nonanalytic term
\begin{equation}
  \Omega_D(m)-\Omega_D(0)
  =
  -K_D\,m^{D-2}
  +O(m^2\Lambda^{D-4}),
  \qquad
  K_D:=
  -\frac{\Gamma(1-D/2)}{(4\pi)^{D/2}}>0.
  \label{eq:on-large-n-omega-expansion}
\end{equation}
The sign is fixed by the fact that \(\Omega_D(m)\) decreases with \(m\).
If \(\tau\) is the linear thermal displacement from the critical saddle, then
\[
  \tau\propto m^{D-2},
  \qquad
  \xi=m^{-1},
\]
and therefore
\begin{equation}
  \nu_{N=\infty}=\frac1{D-2}.
  \label{eq:on-large-n-nu}
\end{equation}
This agrees with the \(N\to\infty\) limit of
\eqref{eq:on-nu} after setting \(D=4-\epsilon\):
\[
  \frac1{D-2}
  =
  \frac1{2-\epsilon}
  =
  \frac12+\frac{\epsilon}{4}+\frac{\epsilon^2}{8}+O(\epsilon^3).
\]

At the critical saddle the vector two-point function is \(1/p^2\), so
\[
  \Delta_\phi=\frac{D-2}{2},
  \qquad
  \eta_{N=\infty}=0.
\]
The singlet thermal operator is represented at large \(N\) by the auxiliary
field \(\sigma\), not by an unrenormalized pointwise product.  Its quadratic
kernel contains the bubble
\begin{equation}
  B_D(p)
  :=
  \int\frac{\dd^D q}{(2\pi)^D}
  \frac1{q^2(q+p)^2}
  =
  b_D\,|p|^{D-4},
  \qquad
  b_D=
  \frac{\Gamma(2-D/2)\Gamma(D/2-1)^2}
       {(4\pi)^{D/2}\Gamma(D-2)} .
  \label{eq:on-large-n-bubble}
\end{equation}
For \(2<D<4\), this term dominates over the local \(U''(\rho_c)\) term at
small momentum.  Hence the \(\sigma\) propagator scales as
\[
  \langle\sigma(p)\sigma(-p)\rangle \propto |p|^{4-D}.
\]
A scalar operator of dimension \(\Delta\) has momentum-space two-point
scaling \(|p|^{2\Delta-D}\).  Thus
\begin{equation}
  \Delta_\sigma=2,
  \qquad
  y_t=D-\Delta_\sigma=D-2,
  \label{eq:on-large-n-sigma-dimension}
\end{equation}
which is the same content as \eqref{eq:on-large-n-nu}.

\begin{proposition}[Large-\(N\) critical saddle]
\label{prop:on-large-n-critical-saddle}
Consider the finite-cutoff \(O(N)\)-invariant scalar model
\eqref{eq:on-large-n-potential-action}, with \(U\) tuned so that the
translation-invariant large-\(N\) saddle has \(m=0\).  Assume \(2<D<4\), the
cutoff is removed only after the saddle equations have been expressed in terms
of a fixed finite critical displacement \(\tau\), and the saddle remains on a
steepest-descent contour obtained from the regulated finite-dimensional
integral.  Then at \(N=\infty\)
\[
  \Delta_\phi=\frac{D-2}{2},
  \qquad
  \eta=0,
  \qquad
  \Delta_\sigma=2,
  \qquad
  \nu=\frac1{D-2}.
\]
\end{proposition}

\begin{proof}
At the critical saddle the inverse vector propagator is \(p^2\), hence the
vector two-point function scales as \(p^{-2}\).  Comparing with the
momentum-space scaling law
\(|p|^{2\Delta_\phi-D}\) gives
\(2\Delta_\phi-D=-2\), or
\(\Delta_\phi=(D-2)/2\), and therefore \(\eta=0\).

The thermal displacement is measured by the saddle mass \(m\).  For
\(2<D<4\), the cutoff integral has the nonanalytic small-\(m\) expansion
\[
  \Omega_D(m)-\Omega_D(0)=-K_Dm^{D-2}+O(m^2\Lambda^{D-4}),
  \qquad K_D>0,
\]
so a finite critical displacement satisfies
\(\tau\propto m^{D-2}\).  Since the correlation length of the vector
propagator is \(\xi=m^{-1}\), this gives
\(\xi\propto |\tau|^{-1/(D-2)}\), hence \(\nu=1/(D-2)\).

Finally, the quadratic kernel of the collective field \(\sigma\) is dominated
in the infrared by the bubble
\[
  B_D(p)=b_D|p|^{D-4}
\]
rather than by the local term.  Its propagator therefore scales as
\(|p|^{4-D}\).  Comparing
\(|p|^{4-D}\) with \(|p|^{2\Delta_\sigma-D}\) gives
\(\Delta_\sigma=2\).  The source dual to \(\sigma\) is the thermal source, so
\(y_t=D-\Delta_\sigma=D-2\), consistently reproducing the value of \(\nu\)
just derived.
\end{proof}

The first nonzero field anomalous dimension is an \(O(1/N)\) effect from
exchange of the collective \(\sigma\) field; the present saddle calculation
establishes the exact \(N=\infty\) dimensions and gives the expansion point
for the systematic \(1/N\) diagrammatics.

Equations \eqref{eq:on-large-n-gap-equations} are the continuum
field-theory form of the spherical model: the fluctuating \(N\)-component spin
length is replaced, at \(N=\infty\), by a saddle constraint on
\(\rho=\phi^2/N\).  This is a statement about the regulated large-\(N\)
limit and its continuum critical saddle, not an independent postulate about
universality.

There is also a lower-critical-dimension chart.  The \(O(N)\) nonlinear sigma
model with field \(n^i(x)\) constrained by \(n^in^i=N\) has regulated action
\begin{equation}
  S_\sigma[n,\alpha]
  =
  \frac1{2T}
  \int\dd^D x\,
  \left[
    \partial_\mu n^i\partial_\mu n^i
    +\alpha(n^in^i-N)
  \right].
  \label{eq:on-nonlinear-sigma-action}
\end{equation}
At \(N=\infty\), integrating over \(N-1\) components gives the same saddle
equations as above, with
\[
  T_c=\Omega_D(0)^{-1},
  \qquad
  \Omega_D(0)-\Omega_D(m)
  =
  K_Dm^{D-2}+\cdots .
\]
Consequently the sigma model has the same large-\(N\) critical exponent
\(\nu=1/(D-2)\) for \(2<D<4\).  Near \(D=2+\widetilde\epsilon\), the usual
dimensionless sigma-model coupling \(\mathfrak g\) obeys, for \(N>2\),
\begin{equation}
  \beta_{\mathfrak g}
  =
  \widetilde\epsilon\,\mathfrak g
  -\frac{N-2}{2\pi}\mathfrak g^2
  +O(\mathfrak g^3),
  \qquad
  \mathfrak g_*
  =
  \frac{2\pi}{N-2}\widetilde\epsilon
  +O(\widetilde\epsilon^2).
  \label{eq:on-nlsm-two-plus-epsilon-beta}
\end{equation}
This fixed point is the \(2+\widetilde\epsilon\) perturbative chart for the
same \(O(N)\) critical family.  The \(N=2\) model is exceptional: the
coefficient \(N-2\) vanishes, and the Berezinskii--Kosterlitz--Thouless
mechanism is not captured by \eqref{eq:on-nlsm-two-plus-epsilon-beta}.

More generally, the Wilson--Fisher epsilon expansion is an asymptotic
small-\(\epsilon\) expansion unless an additional summability theorem is
specified.  Large-order analyses of related perturbative \(\phi^4\) series
exhibit factorial coefficient growth, and no summability theorem is being
assumed here for the displayed epsilon series.  Consequently a finite number
of terms should be read as asymptotic information about the fixed point near
four dimensions.  Extracting accurate \(d=3\) numbers from the epsilon series
requires an extra resummation prescription, for example a Borel or
conformal-Borel method together with assumptions about singularities of the
Borel transform; it is not the same operation as evaluating a convergent Taylor
series at \(\epsilon=1\).

\section{Scaling Coordinates as a Local RG Chart}

The preceding computations identify individual scaling operators.  For the
universality discussion, the corresponding coordinate statement is needed.
Fix a finite projection of the local action space near a fixed point, after
quotienting by total derivatives and equation-of-motion redundancies in the
chosen renormalized-operator scheme.  Let \(g^\alpha(\mu)\) be dimensionless
coordinates in this projection, with \(g^\alpha=0\) at the fixed point, and
let
\[
  \frac{\dd g^\alpha}{\dd\log\mu}
  =
  \beta^\alpha(g)
\]
be the projected RG vector field.  The linearized matrix is
\[
  M^\alpha{}_\beta
  =
  \left.
  \frac{\partial\beta^\alpha}{\partial g^\beta}
  \right|_{g=0}.
\]
Assume, in the finite projection under consideration, that \(M\) is
diagonalizable in a basis represented by scaling operators.  A relevant
coordinate \(u^A\) is one whose operator has dimension
\(\Delta_A<D\); write
\[
  y_A=D-\Delta_A>0 .
\]
An irrelevant coordinate \(v^\rho\) is one whose operator has dimension
\(\Delta_\rho>D\); write
\[
  \omega_\rho=\Delta_\rho-D>0 .
\]
With the convention that \(\mu\) is the running physical momentum scale, the
linearized equations are
\begin{equation}
  \frac{\dd u^A}{\dd\log\mu}
  =
  -y_Au^A+O(g^2),
  \qquad
  \frac{\dd v^\rho}{\dd\log\mu}
  =
  \omega_\rho v^\rho+O(g^2).
  \label{eq:critical-scaling-coordinate-linear-flow}
\end{equation}
Thus a relevant coordinate grows toward the infrared unless it is tuned, while
an irrelevant coordinate decays toward the infrared at the linearized level.
This is a statement about the endpoint map of the RG flow, not about a
microscopic equality of actions.

\subsection{Dangerously Irrelevant Coordinates}
\label{sec:dangerously-irrelevant-coordinates}

Irrelevance is a statement about the RG endpoint in a specified coordinate
chart.  Setting the corresponding coordinate to zero inside an observable is a
separate limiting operation whose validity depends on the physical limit and
on the observable topology.  The precise distinction is as follows.

\begin{definition}[Dangerously irrelevant coordinate for an observable]
Fix a finite RG chart near a fixed point, with relevant coordinates \(u^A\)
and irrelevant coordinates \(v^\rho\) as in
\eqref{eq:critical-scaling-coordinate-linear-flow}.  Let
\({\mathcal O}\) be an observable whose scaling limit is represented in the
same chart by an observable map
\[
  F_{\mathcal O}(u,v)
\]
after the required source normalization and subtraction conventions have been
specified.  An irrelevant coordinate \(v^\rho\) is called dangerously
irrelevant for \({\mathcal O}\), in a sector \(S\) of the relevant-coordinate
space, if the following three conditions hold.

First, \(v^\rho\) is RG-irrelevant at the fixed point:
\(\omega_\rho=\Delta_\rho-D>0\).  Second, for flows approaching the fixed
point through the sector \(S\), the reduced map obtained by first setting
\(v^\rho=0\) is not asymptotic to the full observable map.  Equivalently,
after choosing an infrared rescaling factor \(b\),
\begin{equation}
  F_{\mathcal O}(u,v)
  =
  b^{-\Delta_{\mathcal O}}\,
  {\mathcal F}_{\mathcal O}
  \bigl(b^{y_A}u^A,b^{-\omega_\rho}v^\rho,\ldots\bigr),
  \label{eq:dangerously-irrelevant-scaling-form}
\end{equation}
and the scaling function in the sector under consideration has no finite
limit, or has the wrong limiting branch, as
\(w=b^{-\omega_\rho}v^\rho\to0\).  Third, this singular dependence changes
an exponent or a leading amplitude of \({\mathcal O}\).
\end{definition}

Thus ``dangerous'' is a property of a coordinate together with an observable
and a sector of its scaling variables.  The same coordinate remains irrelevant
for the stability of the fixed point and for observables whose scaling
functions have regular \(w\to0\) limits.

The canonical example is the scalar/Ising equation of state above four
Euclidean dimensions.  At the Gaussian fixed point in \(D>4\), the field has
dimension
\[
  \Delta_\phi=\frac{D-2}{2}.
\]
Let \(M\) denote the uniform order parameter, \(t\) the thermal coordinate,
\(h\) the magnetic source, and \(g>0\) the quartic coordinate in the Landau
functional
\begin{equation}
  f_{\rm L}(M;t,h,g)
  =
  \frac12 tM^2+\frac{g}{4!}M^4-hM .
  \label{eq:dangerous-landau-functional}
\end{equation}
The Gaussian engineering dimensions are
\[
  y_t=2,\qquad
  y_h=D-\Delta_\phi=\frac{D+2}{2},\qquad
  y_g=4-D<0 .
\]
The quartic coordinate is therefore irrelevant at the Gaussian fixed point.
Nevertheless it is load-bearing in the ordered equation of state.  Stationarity
of \eqref{eq:dangerous-landau-functional} gives
\begin{equation}
  h=tM+\frac{g}{6}M^3 .
  \label{eq:dangerous-equation-of-state}
\end{equation}
For \(h=0\) and \(t<0\),
\begin{equation}
  M^2=-\frac{6t}{g}.
  \label{eq:dangerous-ordered-magnetization}
\end{equation}
Setting \(g=0\) before minimizing removes the stabilizing term and destroys
the ordered saddle.  The singular dependence on \(g\) is not a contradiction
of RG irrelevance; it says that the ordered order-parameter observable uses a
singular branch of the scaling function.

To see the exponent directly, apply the Gaussian scaling form
\begin{equation}
  M(t,h,g)
  =
  b^{-(D-2)/2}
  {\mathcal M}
  \left(b^2t,\ b^{(D+2)/2}h,\ b^{4-D}g\right).
  \label{eq:dangerous-magnetization-scaling}
\end{equation}
Choose \(b=|t|^{-1/2}\).  In the ordered sector \(t<0\), \(h=0\), the third
argument is
\[
  w=b^{4-D}g=g\,|t|^{(D-4)/2}\to0^+ .
\]
Equation \eqref{eq:dangerous-ordered-magnetization} implies
\[
  {\mathcal M}(-1,0,w)\sim \left(\frac6w\right)^{1/2}
  \qquad (w\to0^+).
\]
Substitution into \eqref{eq:dangerous-magnetization-scaling} gives
\begin{equation}
  M(t,0,g)
  \sim
  \sqrt{\frac6g}\,|t|^{1/2},
  \qquad D>4,\quad t\to0^- .
  \label{eq:dangerous-mean-field-beta}
\end{equation}
The magnetic critical exponent is therefore the mean-field value
\(\beta_{\rm mag}=1/2\), not the value \((D-2)/4\) obtained by incorrectly
assuming a regular \(w\to0\) limit of
\({\mathcal M}(-1,0,w)\).  At the critical isotherm \(t=0\),
\eqref{eq:dangerous-equation-of-state} gives
\[
  M(0,h,g)\sim\left(\frac{6h}{g}\right)^{1/3},
\]
so \(\delta=3\).  The minimized free-energy density in the ordered phase is
\[
  f_{\rm L,min}(t<0,h=0,g)
  =
  -\,\frac{3t^2}{2g}.
\]
Equivalently, the free-energy scaling function behaves as \(w^{-1}\).  This
is the elementary mechanism behind the failure of naive hyperscaling above
the upper critical dimension: the Gaussian value \(\nu=1/2\) and the
free-energy exponent \(2-\alpha=2\) do not satisfy
\(2-\alpha=D\nu\) when \(D>4\).

At the upper critical dimension \(D=4\), the same coordinate is marginally
irrelevant.  In the one-component normalization
\(x=\lambda/(16\pi^2)\) used in
\eqref{eq:wf-two-loop-beta-x}, put
\[
  L=\log\frac{\mu_0}{\mu}.
\]
The four-dimensional quartic flow satisfies
\begin{equation}
  \frac{\dd x}{\dd L}
  =
  -3x^2+O(x^3),
  \qquad
  x(L)
  =
  \frac{x_0}{1+3x_0L}
  +O(L^{-2})
  \sim\frac1{3L}.
  \label{eq:dangerous-four-dimensional-running-x}
\end{equation}
The thermal coordinate obeys, to the leading logarithmic order needed here,
\begin{equation}
  \frac{\dd t}{\dd L}
  =
  \bigl(2-x(L)+O(x(L)^2)\bigr)t .
  \label{eq:dangerous-four-dimensional-running-t}
\end{equation}
Therefore
\[
  t(L)
  =
  t_0 e^{2L}\bigl(1+3x_0L\bigr)^{-1/3}(1+o(1)).
\]
The correlation-length scale \(L_\xi\) is determined by \(|t(L_\xi)|=O(1)\),
and hence
\begin{equation}
  \xi
  \asymp
  |t_0|^{-1/2}
  \left(\log\frac1{|t_0|}\right)^{1/6}
  \label{eq:dangerous-four-dimensional-xi-log}
\end{equation}
up to a nonuniversal multiplicative constant and lower logarithms.  At that
scale the ordered saddle gives \(M_{\rm scaled}\asymp x(L_\xi)^{-1/2}\), while
the field rescaling contributes \(e^{-L_\xi}\) at leading logarithmic order.
Thus
\begin{equation}
  M(t_0,0,x_0)
  \asymp
  |t_0|^{1/2}
  \left(\log\frac1{|t_0|}\right)^{1/3},
  \qquad t_0\to0^- .
  \label{eq:dangerous-four-dimensional-magnetization-log}
\end{equation}
The logarithm is the marginal version of the singular
\(g^{-1/2}\) dependence in \eqref{eq:dangerous-mean-field-beta}.  The
power exponents are mean-field exponents, while the leading logarithmic
corrections remember the slow RG decay of the quartic coordinate.

\begin{definition}[Critical surface in a finite RG chart]
Let \(u=(u^1,\ldots,u^r)\) be the relevant coordinates in a finite local RG
chart at a fixed point.  At a reference scale \(\mu_R\), the critical surface
in that chart is the set of ultraviolet coordinates whose RG trajectory exists
down to \(\mu_R\) and satisfies
\[
  u(\mu_R)=0 .
\]
If the theory is restricted by a symmetry, only relevant coordinates
compatible with that symmetry are included in \(u\).
\end{definition}

This definition is set-theoretic.  Its promotion to a \(C^k\) submanifold
statement requires the finite-reference implicit-function construction: one
must specify a Banach or finite-dimensional RG chart, prove that the endpoint
map is \(C^k\), and prove that its derivative in the relevant directions is
invertible.  The linearized eigenvalue calculation identifies the candidate
relevant coordinates; it does not by itself prove smoothness or codimension of
the nonlinear tuned set.

At linear order between a microscopic cutoff \(\Lambda_0\) and a reference
scale \(\mu_R<\Lambda_0\), equation
\eqref{eq:critical-scaling-coordinate-linear-flow} gives
\[
  u^A(\mu_R)
  =
  \left(\frac{\Lambda_0}{\mu_R}\right)^{y_A}
  u^A(\Lambda_0),
  \qquad
  v^\rho(\mu_R)
  =
  \left(\frac{\mu_R}{\Lambda_0}\right)^{\omega_\rho}
  v^\rho(\Lambda_0).
\]
Therefore holding \(u^A(\mu_R)\) fixed as \(\Lambda_0\to\infty\) requires
the ultraviolet coordinate \(u^A(\Lambda_0)\) to be tuned with precision of
order \((\mu_R/\Lambda_0)^{y_A}\).  A bounded ultraviolet irrelevant coordinate
contributes to \(v^\rho(\mu_R)\) only at order
\((\mu_R/\Lambda_0)^{\omega_\rho}\), apart from the generated contribution
along the tuned trajectory.  The Wilsonian cutoff-removal estimate of
Chapter~\ref{ch:vol2-wilsonian-effective-actions} gives the nonlinear
finite-coordinate form of this statement.

For the Wilson-Fisher fixed point in the \(\mathbb Z_2\)-even scalar chart,
the leading relevant coordinate is the thermal coordinate dual to
\([O_2]\), with \(y_t=D-\Delta_{\phi^2}\).  In the full scalar chart the odd
coordinate dual to \(\phi\) is also relevant, with
\(y_h=D-\Delta_\phi\).  The phrase ``codimension-one critical surface'' will
therefore mean the \(\mathbb Z_2\)-even chart and the finite-reference
submanifold theorem above, unless the magnetic coordinate has been included
and separately tuned.

Chapter~\ref{ch:ising-universality} applies this fixed-point data to the
statistical Ising model.  The coarse-grained Ising variables determine which
local scalar deformations are compatible with spin-flip symmetry, and the
Wilson-Fisher eigenvalues \(y_t\), \(y_h\), and \(\omega_\rho\), together with
the exponents \(\eta\) and \(\nu\), determine the scaling of the corresponding
thermal, magnetic, and irrelevant directions near criticality.
