\chapter{Instantons as Physical Amplitudes}
\label{ch:instantons-physical-amplitudes}

Chapter~\ref{ch:volume-ii-20} developed the BPST solution, the
finite-regulator instanton density, the anomaly selection rules, and several
QCD amplitude benchmarks.  This chapter reorganizes those ingredients around
the physical object they are meant to compute.  The object is not the
instanton moduli space and not the universal density by itself.  It is a
regulated source or state amplitude, with zero modes saturated, nonzero
fluctuations divided by the chosen counterterms, endpoint faces controlled,
and a Lorentzian or Euclidean observable projection specified.

The distinction is operational.  Two channels may use exactly the same BPST
collective-coordinate space and the same one-loop density, while one channel
vanishes because its sources have rank one in the zero-mode space and the
other gives a nonzero four-fermion coefficient.  Or two Euclidean source
coefficients may agree after integrating over the collective variables, while
their spectral projections differ.  These are not refinements of moduli-space
geometry; they are the steps which turn a saddle contribution into a QFT
amplitude.

The chapter follows the order of that calculation.  The first object is the
finite-regulator source functional: before an amplitude exists, the source
derivatives, zero-mode Berezin coefficient, normal-fluctuation average, and
collective integral must all be in the same channel.  The density
normalization is then separated into collective coordinates, zero-mode
normalizations, nonzero-mode determinants, running-coupling conversion, and
scheme constants.  The channel supplies source ranks, amputation, endpoint
windows, and a declared physical projection.  The hard \(SU(3)\), \(N_f=2\)
benchmark then shows how zero-mode form factors, tail subtraction, and a
Wilsonian size split enter one concrete coefficient.  Finally,
normal-fluctuation source response, cluster corrections, and observable maps
decide which part of the instanton ensemble is being measured.

\section{The Source Functional Before the Amplitude}
\label{sec:instanton-source-functional-route}

At finite regulator the one-instanton sector first produces a source
functional.  A physical amplitude is a later extraction from that functional.
This order is the main guard against confusing an instanton measure with an
observable.  In a retained one-instanton chart \(z\), after collective and
gauge zero modes have been separated, write the fermion zero-mode
coordinates as \(\alpha,\bar\alpha\) and the nonzero-mode fluctuation as
\(\eta\).  A typical retained contribution has the order-sensitive finite
form
\begin{equation}
\begin{aligned}
  Z_{\Lambda}^{I}[J]
  &=
  \ee^{\ii\theta}
  \int_{\mathcal U_{\rm win}}
  \ee^{-S_{\Lambda}(A_z)}
  \mathcal W_{\Lambda}^{\rm nz}(z)\,
  \mathcal B_{\Lambda}^{\rho}(z)                         \\
  &\qquad{}\times
  \left\langle
    \int
    \dd\bar\alpha\,\dd\alpha\,
    \exp\!\left[
      \bar\alpha
      \bigl(M_z+\mathcal S_z[J,\eta]\bigr)
      \alpha
    \right]\,
    \mathcal I_z[J,\eta]
    \ee^{-V_z(\eta)}
  \right\rangle_{\gamma_z}
  \dd\nu_{\Lambda,\mathcal M}(z).
\end{aligned}
  \label{eq:instanton-source-functional-route}
\end{equation}
Here \(\dd\nu_{\Lambda,\mathcal M}\) is the collective-coordinate density,
\(\mathcal W_{\Lambda}^{\rm nz}\) is the zero-mode-deleted determinant
normalization, \(\mathcal B_{\Lambda}^{\rho}\) records the size-window or
endpoint factor, \(M_z+\mathcal S_z[J,\eta]\) is the restriction of masses and
sources to the fermion zero-mode space, \(\mathcal I_z[J,\eta]\) is the
remaining source insertion, and \(V_z\) is the non-Gaussian fluctuation
remainder.  The displayed brackets are the Gaussian normal-mode average; the
source dependence inside the same average is not fixed by the determinant
constant alone.

A selected observable is obtained only after applying the appropriate source
differential and physical projection:
\begin{equation}
  \mathcal A_{\mathcal O,\Lambda}^{I}
  =
  \Pi_{\rm phys}^{\mathcal O}
  \left[
    \mathsf D_{\mathcal O,J}
    Z_{\Lambda}^{I}[J]\big|_{J=0}
  \right]
  +R_{\mathcal O,\Lambda}.
  \label{eq:instanton-source-functional-observable-extraction}
\end{equation}
The operator \(\mathsf D_{\mathcal O,J}\) may be a Grassmann source
differentiation, a gauge-invariant source matching map, or an OPE
coefficient extraction.  The projection
\(\Pi_{\rm phys}^{\mathcal O}\) may be a pole residue, a spectral bin, an
inclusive cut, or a Euclidean susceptibility functional.  The residual
\(R_{\mathcal O,\Lambda}\) contains precisely the regulator, continuation,
sector, endpoint, and matching errors that are invisible in the moduli
measure.

\begin{controlledapproximation}[Source-functional route to a physical amplitude]
\label{ca:instanton-source-functional-route}
The route
\[
\begin{aligned}
  &\hbox{saddle}
  \longrightarrow
  \hbox{zero-mode source functional}                         \\
  &\longrightarrow
  \hbox{normal-fluctuation source average}
  \longrightarrow
  \hbox{collective-coordinate integral}                       \\
  &\longrightarrow
  \hbox{physical projection}
\end{aligned}
\]
cannot in general be reordered.  Three finite checks already show where the
shortcuts fail.

First, source differentiation is not mass saturation.  For two flavors,
\[
  \det(M+B)
  =
  m_um_d+m_uB_{dd}+m_dB_{uu}+B_{uu}B_{dd}-B_{ud}B_{du},
\]
so
\[
  \partial_{B_{uu}}\partial_{B_{dd}}\det(M+B)\big|_{B=0}=1,
  \qquad
  \partial_{B_{ud}}\partial_{B_{du}}\det(M+B)\big|_{B=0}=-1,
\]
whereas the vacuum activity is \(m_um_d\).  The differentiated
four-source coefficient is therefore a different coordinate of the
zero-mode determinant, not a reinterpretation of the mass-saturated term.

Second, determinant normalization is not source response.  Even after
\(\mathcal W_{\Lambda}^{\rm nz}\) has fixed the Gaussian determinant,
\[
  \frac{\langle\mathcal I_J(\eta)\ee^{-V(\eta)}\rangle_\gamma}
       {\langle\ee^{-V(\eta)}\rangle_\gamma}
  =
  \langle\mathcal I_J\rangle_\gamma
  +
  \frac{
    \operatorname{Cov}_\gamma(\mathcal I_J,\ee^{-V})
  }{
    \langle\ee^{-V}\rangle_\gamma
  } .
\]
The covariance term is source-dependent amplitude data.  It vanishes only by
a separate fluctuation argument, not by quoting the determinant constant.

Third, Euclidean source kernels are not physical observables before
projection.  A contact polynomial can change a spacelike Euclidean source
value while contributing no separated spectral-bin weight, and a colored
partonic kernel in a confining theory has no standalone LSZ residue.  Thus
the object in \eqref{eq:instanton-source-functional-route} is the input to
an observable map; \eqref{eq:instanton-source-functional-observable-extraction}
is the physical claim.
\end{controlledapproximation}

\begin{controlledapproximation}[Finite source-functional amplitude laboratory]
\label{ca:instanton-finite-source-functional-laboratory}
A finite retained-window laboratory makes the same order visible in one
amplitude coordinate.  Let \(c\) run through finitely many retained
collective cells.  In cell \(c\), write \(\kappa_c\) for the collective
weight, \(\Delta_c^{\rm nz}\) for the zero-mode-deleted determinant
normalization, \(z_c\) for the differentiated zero-mode Berezin coordinate,
and \(p_c\) for the declared physical projection.  Let the normal fluctuation
average be represented by two finite normal points \(\sigma=\pm\), with
positive interaction weights \(r_{c\sigma}\) and source insertion values
\(u_{c\sigma}\).  Define
\[
  D_c^{\rm nor}:=\frac{1}{2}\sum_{\sigma=\pm}r_{c\sigma},
  \qquad
  \bar U_c:=\frac{1}{2}\sum_{\sigma=\pm}u_{c\sigma},
  \qquad
  R_c^{\rm src}:=
  \frac{
    \sum_{\sigma=\pm}u_{c\sigma}r_{c\sigma}
  }{
    \sum_{\sigma=\pm}r_{c\sigma}
  } .
\]
Equivalently,
\[
  R_c^{\rm src}
  =
  \bar U_c+
  \frac{\operatorname{Cov}_c(U,r)}{D_c^{\rm nor}},
  \qquad
  \operatorname{Cov}_c(U,r)
  :=
  \frac{1}{2}\sum_{\sigma=\pm}
  (u_{c\sigma}-\bar U_c)(r_{c\sigma}-D_c^{\rm nor}).
\]
The finite amplitude laboratory is then
\begin{equation}
  A_{\rm lab}
  =
  \sum_c
  p_c\,\kappa_c\,\Delta_c^{\rm nz}\,z_c\,R_c^{\rm src}.
  \label{eq:instanton-finite-source-functional-laboratory}
\end{equation}
The five factors have different physical origins: collective measure,
nonzero-mode determinant, zero-mode source differentiation, normal-mode
source response, and observable projection.  Replacing \(z_c\) by the
mass-saturated activity \(m_um_d\), replacing \(R_c^{\rm src}\) by the
Gaussian mean \(\bar U_c\), or dropping \(p_c\) gives respectively
\[
  A_{\rm mass},\qquad
  A_{\rm det\text{-}only},\qquad
  A_E,
\]
which are generally different finite numbers.  The laboratory is not a
continuum proof of an instanton amplitude.  Its role is stricter and more
basic: a calculation that supplies only an ADHM or BPST collective volume has
at most \(\kappa_c\), and perhaps \(\Delta_c^{\rm nz}\).  It has not yet
computed the physical amplitude until the source determinant coordinate,
normal-fluctuation response, and projection covector have also been supplied
in the same channel.
\end{controlledapproximation}

\section{The Density Normalization: Fluctuations Before Channels}
\label{sec:instanton-density-gate}

The density in \eqref{eq:one-instanton-universal-density} is already a
fluctuation calculation.  The BPST collective-coordinate volume form supplies
one of its factors.  In the half-trace convention of
Chapter~\ref{ch:volume-ii-20}, its universal one-loop part is
\begin{equation}
  \dd\mu_1^{\rm univ}
  =
  C_{\mathcal S,N_c,N_f}\,
  \frac{\dd^4a\,\dd\rho}{\rho^5}\,
  \dd\Omega_{N_c}\,
  \left(\frac{8\pi^2}{g_{\rm ht}^2(\mu)}\right)^{2N_c}
  (\mu\rho)^{b_0}
  \exp\!\left[-\frac{8\pi^2}{g_{\rm ht}^2(\mu)}+\ii\theta\right]
  \left(1+O(g_{\rm ht}^2)\right),
  \label{eq:instanton-physical-density-gate-universal}
\end{equation}
with \(b_0\) defined in
\eqref{eq:instanton-density-beta0-definition}.  The factors have different
origins.  The \(\dd^4a\,\dd\rho\,\rho^{-5}\dd\Omega_{N_c}\) part is the
collective-coordinate Jacobian and orientation convention.  The power
\((8\pi^2/g_{\rm ht}^2)^{2N_c}\) comes from the \(4N_c\) bosonic zero-mode
normalizations.  The exponential is the classical BPST action.  The factor
\((\mu\rho)^{b_0}\) is the logarithmic remnant of the zero-mode-deleted
gauge, ghost, and matter fluctuation determinants after counterterm
subtraction.  The finite constant \(C_{\mathcal S,N_c,N_f}\) is not fixed by
dimensional analysis; it is part of the chosen determinant and orientation
normalization scheme.

\begin{controlledapproximation}[One-loop density normalization for a channel]
\label{ca:instanton-one-loop-density-gate-channel}
The one-loop scale test which fixes the logarithmic density power is local:
\[
  \left(\mu\partial_\mu+\beta(g_{\rm ht})\partial_{g_{\rm ht}}\right)
  \log\!\left[
    (\mu\rho)^\alpha
    \exp\!\left(-\frac{8\pi^2}{g_{\rm ht}^2(\mu)}\right)
  \right]
  =
  \alpha-b_0+O(g_{\rm ht}^2).
\]
Thus the density is RG consistent only for \(\alpha=b_0\).  This test
determines the logarithmic fluctuation power, not the full physical
amplitude.  Once the density is inserted into a channel, the retained
integrand has the channel form
\begin{equation}
  \dd\mathcal A_{\mathcal C}^{\rm win}
  =
  \dd\rho\,
  \rho^{b_0-5}\,
  \mathcal W_{\rm fin}^{\rm nz}(\rho,U)\,
  \mathcal Z_{\mathcal C}^{0}(\rho,U)\,
  \mathcal M_{\mathcal C}^{\rm src}(\rho,U)\,
  \mathcal B_{\mathcal C}^{\rho}(\rho)\,
  \dd^4a\,\dd\Omega_{N_c}
  \times(\hbox{scheme and action factors}).
  \label{eq:instanton-density-gate-channel-integrand}
\end{equation}
If the zero-mode/source part starts as
\[
  \mathcal Z_{\mathcal C}^{0}\mathcal M_{\mathcal C}^{\rm src}
  =
  \rho^{\beta_{\mathcal C}}
  \bigl(Z_{\mathcal C}^{(0)}+O(\rho^\delta)\bigr),
\]
then the channel size power is
\(\rho^{b_0+\beta_{\mathcal C}-5}\dd\rho\).  The number
\(\beta_{\mathcal C}\) is physical source data.  It changes when masses
saturate zero modes, when external source derivatives are taken, when
amputation conventions are changed, or when the observable projection kills a
zero-mode component.  For the hard \(SU(3)\), \(N_f=2\), four-source channel
below, \(\beta_{\mathcal C}=6\), so the integrand begins as
\(\rho^{32/3}\dd\rho\), not as the density-only power
\(\rho^{14/3}\dd\rho\).

Consequently a same-theory ratio may cancel the finite determinant constant
and \(\Lambda_{\rm ht}^{b_0}\) only after the source frame, zero-mode
determinants, orientation projection, endpoint window, and physical
projection have been transported.  A computation which keeps
\eqref{eq:instanton-physical-density-gate-universal} but drops
\(\mathcal Z_{\mathcal C}^{0}\), \(\mathcal M_{\mathcal C}^{\rm src}\), or
\(\mathcal B_{\mathcal C}^{\rho}\) has not computed the channel amplitude.
\end{controlledapproximation}

\begin{controlledapproximation}[Zero-mode Jacobian from a gauge slice]
\label{ca:instanton-collective-jacobian-gauge-slice}
The collective measure is also a fluctuation calculation, not a dimension
count.  Choose a finite regulator and a local slice through the instanton
orbit.  For coordinates \(q^i\) on the retained chart, let
\[
  \psi_i(z)
  =
  \Pi_z^{\perp}
  \frac{\partial A_z}{\partial q^i},
  \qquad i=1,\ldots,m,
\]
be the horizontal zero-mode representatives after infinitesimal gauge
directions and any residual stabilizer have been removed.  The bosonic
zero-mode metric in the same action and counterterm convention is
\begin{equation}
  \mathcal G_{ij}^{0}(z)
  =
  \frac{1}{g_{\rm ht}^2(\mu)}
  \langle \psi_i(z),\psi_j(z)\rangle_{\Lambda}
  +\mathcal G^{\rm ct}_{ij}(z).
  \label{eq:instanton-zero-mode-gram-metric}
\end{equation}
If the path-integral measure is written in orthonormal zero-mode
coefficients \(c_r\), the zero-mode replacement is
\begin{equation}
  \prod_{r=1}^{m}\frac{\dd c_r}{\sqrt{2\pi}}
  \quad\rightsquigarrow\quad
  \frac{\sqrt{\det\mathcal G^{0}(z)}}{(2\pi)^{m/2}\mathcal V_{\rm stab}(z)}
  \prod_{i=1}^{m}\dd q^i .
  \label{eq:instanton-zero-mode-jacobian-replacement}
\end{equation}
Here \(\mathcal V_{\rm stab}\) is the finite residual gauge stabilizer or
quotient volume in the chosen slice.  For a charge-one \(SU(N_c)\) BPST
channel this is the origin of the bosonic factor
\[
  \frac{\dd^4a\,\dd\rho}{\rho^5}\,
  \dd\Omega_{N_c}\,
  \left(\frac{8\pi^2}{g_{\rm ht}^2(\mu)}\right)^{2N_c}
  \mathcal J_{\rm fin}^{0}(\rho,\Omega),
\]
before the zero-mode-deleted determinant and the source channel are
inserted.  The number \(4N_c\) fixes only how many zero-mode factors appear;
it does not determine the \(\rho^{-5}\) density, the orientation quotient, or
the finite normalization.

This replacement is a physical-amplitude normalization because it multiplies
the same source kernel that is later amputated and projected.  Let
\(\mathcal G=\mathcal G^{0}+\delta\mathcal G\) be the metric actually used in
the retained regulator window, and set
\[
  E=(\mathcal G^{0})^{-1}\delta\mathcal G,\qquad
  \|E\|\le r<1
\]
for any submultiplicative matrix norm.  Then
\begin{equation}
  \left|
  \frac{\sqrt{\det\mathcal G}}{\sqrt{\det\mathcal G^{0}}}
  -1
  \right|
  \le
  \exp\!\left(\frac{mr}{2(1-r)}\right)-1 .
  \label{eq:instanton-zero-mode-jacobian-stability}
\end{equation}
Thus a same-density statement must transport the gauge slice, the Gram
determinant, the residual stabilizer, and the source frame.  Using raw
\(\partial_iA_z\) vectors before horizontal projection includes pure-gauge
directions; replacing \(\sqrt{\det\mathcal G}\) by \(\det\mathcal G\) gives
the wrong functional-measure power; and counting zero modes cannot determine
whether the selected fermion source determinant or physical projection
vanishes.
\end{controlledapproximation}

\begin{controlledapproximation}[Proper-time determinant log inside a source
channel]
\label{ca:instanton-proper-time-determinant-channel}
The determinant power in
\eqref{eq:instanton-physical-density-gate-universal} is obtained before any
particular source is differentiated.  At a retained chart point \(z\), after
collective and gauge zero modes have been projected out, write the
zero-mode-deleted one-loop factor as
\begin{equation}
\begin{aligned}
  \log\mathcal W^{\rm nz}_{\Lambda}(z)
  &=
  -\frac12\operatorname{Tr}'\log
  \frac{\Delta_{B,z}}{\Delta_{B,0}}
  +\operatorname{Tr}'\log
  \frac{M_{{\rm gh},z}}{M_{{\rm gh},0}}                  \\
  &\qquad
  +\sum_f\log\det'
  \frac{\mathcal D_{f,z}}{\mathcal D_{f,0}}
  -S^{\rm ct}_{\Lambda}(z).
\end{aligned}
  \label{eq:instanton-proper-time-log-determinant-channel}
\end{equation}
For Dirac fundamental fermions this is a determinant.  A Pfaffian notation
would require an explicitly antisymmetric Majorana, or doubled-field,
coordinate with the corresponding one-half bookkeeping; it is not the
benchmark convention used here.
Each logarithm is defined by a proper-time difference, for example
\[
  \operatorname{Tr}'\log\frac{A_z}{A_0}
  =
  -\int_{\varepsilon}^{T}
  \frac{\dd t}{t}\,
  \operatorname{Tr}'\!\left(\ee^{-tA_z}-\ee^{-tA_0}\right)
  +\hbox{finite subtraction}.
\]
If the logarithmic small-\(t\) coefficients in the boson, ghost, and fermion
blocks are \(h_B,h_{\rm gh},h_F\), and the local counterterm/coupling
conversion contributes \(h_{\rm ct}\) in the same half-trace convention, then
the retained logarithmic power is
\begin{equation}
  \alpha_{\rm det}
  =
  -\frac12 h_B+h_{\rm gh}+h_F+h_{\rm ct}.
  \label{eq:instanton-proper-time-channel-log-power}
\end{equation}
For a one-instanton density this number must equal \(b_0\), so that the
proper-time determinant logarithm cancels the one-loop running of the BPST
action.  In the \(SU(3)\), \(N_f=2\) hard benchmark below, one convenient
split has nonzero-mode spectral contribution \(7/3\) and local
counterterm/coupling contribution \(22/3\), giving \(29/3=b_0\).

This determinant calculation still does not define a four-source amplitude.
Let \(K_{\mathcal C,\Lambda}^{(0)}(z)\) denote the channel kernel obtained
from the leading determinant power, the collective density, the zero-mode
Berezin coefficient, the source/amputation map, and the endpoint factor.  If
the remaining determinant logarithm satisfies
\(|\delta_{\rm det}(z)|\le\epsilon_{\rm det}(z)\) on the retained window, then
\begin{equation}
  \left|
  \mathcal A_{\mathcal C,\Lambda}^{\rm exact}
  -
  \mathcal A_{\mathcal C,\Lambda}^{(0)}
  \right|
  \le
  \int_{\mathcal U_{\rm win}}
  |K_{\mathcal C,\Lambda}^{(0)}(z)|
  \bigl(\ee^{\epsilon_{\rm det}(z)}-1\bigr)
  \dd\nu_{\Lambda,\mathcal M}(z).
  \label{eq:instanton-proper-time-channel-window-bound}
\end{equation}
The absolute value is essential: signed cancellations among heat-kernel
remainders do not control a physical source window.  Also, if the
zero-mode/source determinant or the physical projection vanishes, a nonzero
\(\mathcal W^{\rm nz}_{\Lambda}\) only proves that the fluctuation density is
nonzero; it does not revive the channel amplitude.
\end{controlledapproximation}

\section{The Channel Is Part of the Calculation}
\label{sec:instanton-physical-amplitude-channel}

The finite-regulator inputs
of Definition~\ref{def:one-instanton-amplitude-datum} are the local starting
point for the present chapter.  We now package them as an observable channel.
This is a change of emphasis, not a new convention: the BPST/ADHM data select
the domain of integration, while the channel records the determinant,
zero-mode, source, endpoint, and physical-projection data which decide the
amplitude.

\begin{definition}[Instanton physical-amplitude channel]
\label{def:instanton-physical-amplitude-channel}
A one-instanton physical-amplitude channel at regulator \(\Lambda\) is a
tuple
\begin{equation}
  \mathcal C_{\Lambda}
  =
  \bigl(
  \mathcal U_{\rm win},
  \dd\nu_{\Lambda,\mathcal M},
  \mathcal W_{\Lambda}^{\rm nz},
  \mathcal Z_{\Lambda}^{0},
  \mathcal M_{\Lambda}^{\rm src},
  \mathcal B_{\Lambda}^{\rho},
  \mathfrak L_{\rm phys}
  \bigr).
  \label{eq:instanton-physical-amplitude-channel}
\end{equation}
Here \(\mathcal U_{\rm win}\) is the retained chart/window in center, size,
and orientation variables; \(\dd\nu_{\Lambda,\mathcal M}\) is the
collective-coordinate density; \(\mathcal W_{\Lambda}^{\rm nz}\) is the
zero-mode-deleted fluctuation determinant with its counterterm convention;
\(\mathcal Z_{\Lambda}^{0}\) is the Berezin coefficient which saturates the
fermion zero modes; \(\mathcal M_{\Lambda}^{\rm src}\) is the finite source,
amputation, OPE, color-singlet, or hadron-overlap map;
\(\mathcal B_{\Lambda}^{\rho}\) is the endpoint, size-window, or screening
factor; and \(\mathfrak L_{\rm phys}\) is the final observable projection,
such as an LSZ residue, spectral-bin functional, Euclidean susceptibility,
OPE coefficient, or inclusive cut.
\end{definition}

The channel coefficient is
\begin{equation}
  \mathcal A_{\Lambda}(\mathcal C)
  =
  \mathfrak L_{\rm phys}
  \left[
  \int_{\mathcal U_{\rm win}}
  \ee^{-S_{\Lambda}(A_z)+\ii\theta}\,
  \mathcal W_{\Lambda}^{\rm nz}(z)\,
  \mathcal Z_{\Lambda}^{0}(z)\,
  \mathcal M_{\Lambda}^{\rm src}(z)\,
  \mathcal B_{\Lambda}^{\rho}(z)\,
  \dd\nu_{\Lambda,\mathcal M}(z)
  \right].
  \label{eq:instanton-physical-channel-functional}
\end{equation}
The physical statement is obtained only with a residual budget:
\begin{equation}
\begin{aligned}
  \mathcal A_{\rm phys}(\mathcal C)
  &=
  \mathcal A_{\Lambda}(\mathcal C)
  +R_{\rm det}+R_{\rm fluc}+R_{\rm zm}+R_{\rm src}+R_{\rm Schur}
   +R_{\rho}                                      \\
  &\qquad
  +R_{\rm sec}+R_{\rm cont}+R_{\rm spec}
   +R_{\rm IR}+R_{\rm cut}+R_{\rm scheme}.
\end{aligned}
  \label{eq:instanton-physical-channel-residual-budget}
\end{equation}
The names match the reusable slots in
Proposition~\ref{prop:instanton-amplitude-assembly-error-budget}.  A
calculation may set some of them to zero only by giving the corresponding
argument; a local BPST density does not set any of the source, endpoint, or
physical-projection terms to zero.

\begin{proposition}[Moduli-equivalent channels need not have the same amplitude]
\label{prop:instanton-moduli-equivalent-channel-separation}
Consider two channels with the same finite instanton chart, the same
collective-coordinate density, and the same nonzero-mode determinant
convention.  In a two-flavor zero-mode sector suppose that the only
channel-dependent local source input is a \(2\times2\) zero-mode source
matrix \(B(z)\), so that
\begin{equation}
  \mathcal Z_{\Lambda}^{0}(z)=
  \det B(z)
  =
  B_{uu}(z)B_{dd}(z)-B_{ud}(z)B_{du}(z).
  \label{eq:instanton-two-flavor-source-determinant}
\end{equation}
Then the common moduli measure and determinant convention do not determine
the amplitude.  If \(B(z)\) has rank one on the retained window, the
two-flavor source contribution vanishes identically, even when the
collective-coordinate density and nonzero-mode determinant are nonzero.  If
another channel has a full-rank \(B'(z)\) on a set of positive retained
weight, its coefficient in
\eqref{eq:instanton-physical-channel-functional} can be nonzero with the same
moduli data.
\end{proposition}

\begin{proof}
At finite regulator, the fermion zero modes are ordinary Grassmann variables.
For two flavors the Berezin integral over the four source variables extracts
the determinant coefficient in
\eqref{eq:instanton-two-flavor-source-determinant}.  A rank-one source matrix
has zero determinant pointwise, so the whole zero-mode coefficient is zero
before the collective-coordinate integral is performed.  A full-rank source
matrix has nonzero determinant at the point under consideration.  Multiplying
by the same bosonic density and the same nonzero-mode determinant therefore
changes the channel coefficient only through the source determinant and the
subsequent observable projection.  The common moduli space has not changed.
\end{proof}

This proposition is the shortest way to state the physics lesson.  The
instanton solution supplies a saddle and a measure factor.  The physical
amplitude asks which external states or sources can see the zero-mode
directions, whether that source information survives projection, and whether
the remaining integral is controlled at its boundaries.

\section{Mass Saturation Versus Source Differentiation}
\label{sec:instanton-mass-source-differentiation}

The local 't Hooft determinant is useful precisely because it keeps the
zero-mode source algebra visible.  It becomes misleading when the determinant
is read as a single undifferentiated number, independent of how the physical
observable is selected.

\begin{proposition}[Two-flavor mass/source determinant coordinate]
\label{prop:two-flavor-source-mass-determinant-coordinate}
Let
\[
  M=\begin{pmatrix}m_u&0\\0&m_d\end{pmatrix},
  \qquad
  B=\begin{pmatrix}B_{uu}&B_{ud}\\B_{du}&B_{dd}\end{pmatrix}
\]
be the mass and source restrictions to the instanton zero-mode space.  The
finite zero-mode Berezin coefficient is
\begin{equation}
  \det(M+B)
  =
  m_um_d
  +m_uB_{dd}+m_dB_{uu}
  +B_{uu}B_{dd}-B_{ud}B_{du}.
  \label{eq:two-flavor-mass-source-determinant-polynomial}
\end{equation}
Thus the mass-saturated vacuum activity, the two-source mass-assisted
coefficients, and the four-source differentiated coefficient are different
coordinates of the same zero-mode determinant.  They are not the same
physical observable.
\end{proposition}

\begin{proof}
This is the determinant of a \(2\times2\) matrix:
\[
  \det(M+B)
  =(m_u+B_{uu})(m_d+B_{dd})-B_{ud}B_{du}.
\]
Expanding gives
\eqref{eq:two-flavor-mass-source-determinant-polynomial}.  Berezin
integration then selects the term with the required number of source
variables or mass insertions.  Setting \(B=0\) leaves the vacuum coordinate
\(m_um_d\); setting \(M=0\) leaves the four-source coefficient
\(\det B\).  Mixed mass/source observables select the linear-in-\(B\)
terms.
\end{proof}

This separation is what the original instanton amplitude calculation uses.
The four-fermion coefficient is obtained by differentiating the source
functional and retaining the zero-mode form factors, not by quoting the
mass-saturated vacuum activity and changing its interpretation.  The
coefficient must then be paired with the nonzero-mode determinant, the color
orientation tensor, external source residues or LSZ data, and a size-window
or infrared mechanism.

\begin{proposition}[Chirality-source selection for the instanton vertex]
\label{prop:instanton-chirality-source-selection-gate}
Fix the \(Q=1\) convention in which the fermion zero-mode source block is
\(B^{(+)}\), coupling one left-handed quark source slot to one right-handed
antiquark source slot for each flavor.  If \(\alpha_f,\bar\alpha_f\) are the
zero-mode Grassmann coefficients, then
\begin{equation}
  Z_{\rm zm}^{(+)}[B]
  =
  \int
  \prod_{f=1}^{N_f}\dd\bar\alpha_f\,\dd\alpha_f\,
  \exp\!\left(
    \sum_{f,g=1}^{N_f}\bar\alpha_f B^{(+)}_{fg}\alpha_g
  \right)
  =
  \det B^{(+)} .
  \label{eq:instanton-chirality-source-gate-determinant}
\end{equation}
The differentiated hard vertex in this sector is therefore supported only on
the source coordinate with \(N_f\) left-handed quark slots and \(N_f\)
right-handed antiquark slots, up to replacements of matched pairs by mass
insertions.  In source-slot notation,
\[
  (n_L,n_R,\bar n_L,\bar n_R)=(N_f,0,0,N_f)
\]
for the massless \(Q=1\) hard coefficient, while the \(Q=-1\) coefficient uses
the conjugate coordinate \((0,N_f,N_f,0)\).  The \(Q=1\) zero-mode vertex has
axial weight \(-2N_f\) in this convention, and the anti-instanton vertex has
the opposite weight.
\end{proposition}

\begin{proof}
At finite regulator the zero-mode variables are a finite Grassmann algebra.
The Berezin integral is nonzero only when the expansion of the exponential
contains every \(\alpha_f\) and every \(\bar\alpha_f\) exactly once.  Expanding
the exponential in \eqref{eq:instanton-chirality-source-gate-determinant}
therefore gives the determinant of the source matrix \(B^{(+)}\).  A source
derivative in the conjugate chirality block has no \(Q=1\) zero-mode variable
to saturate, and a chirality-balanced four-source derivative leaves at least
one of the \(Q=1\) zero-mode coefficients unsaturated.  Such a channel may
exist perturbatively or in the anti-instanton sector, but it is not the
\(Q=1\) 't Hooft zero-mode vertex.  The axial weight follows from
\(q_L\mapsto e^{-\ii a}q_L\), \(\bar q_R\mapsto e^{-\ii a}\bar q_R\), so each
matched \(Q=1\) zero-mode source pair carries weight \(-2\).
\end{proof}

For \(N_f=2\), this proposition is the chirality-flow version of
\eqref{eq:two-flavor-mass-source-determinant-polynomial}.  A nonzero
determinant in the wrong chirality block, or a chirality-balanced external
four-fermion source, cannot be moved into the one-instanton hard coefficient
by multiplying by the BPST density.  The leading instanton amplitude first
violates the anomalous axial charge in the zero-mode Berezin integral, then
receives the determinant, color-orientation, normal-fluctuation, endpoint, and
physical-projection factors.  Reversing that order is a different physical
calculation.

\begin{proposition}[Axial Ward transport of the source determinant]
\label{prop:instanton-axial-ward-source-transport}
For \(N_f=2\), keep the \(Q=1\) source convention of
Proposition~\ref{prop:instanton-chirality-source-selection-gate}, and write the
zero-mode mass/source matrix as
\[
  C_{\rm zm}=S+\ii P ,
\]
where \(S\) and \(P\) are the scalar and pseudoscalar source matrices in the
same finite-regulator source frame as
Proposition~\ref{prop:two-flavor-instanton-source-ward-ledger}.  The singlet
axial Ward vector acts on these source coordinates by
\[
  \delta_\alpha S=2\alpha P,\qquad
  \delta_\alpha P=-2\alpha S,\qquad
  \delta_\alpha\theta=4\alpha .
\]
Equivalently,
\[
  \mathcal V_A^{\rm src}C_{\rm zm}=-2\ii C_{\rm zm},
  \qquad
  \mathcal V_A\theta=4 .
\]
Then the \(Q=1\) zero-mode source factor is Ward-transported together with the
instanton phase:
\begin{equation}
  \left(4\partial_\theta+\mathcal V_A^{\rm src}\right)
  \left(\ee^{\ii\theta}\det C_{\rm zm}\right)=0 .
  \label{eq:instanton-source-determinant-ward-transport}
\end{equation}
The anti-instanton sector obeys the conjugate identity with
\(\ee^{-\ii\theta}\det C_{\rm zm}^{\dagger}\).
\end{proposition}

\begin{proof}
The source variation gives
\[
  \delta_\alpha C_{\rm zm}
  =
  2\alpha P-2\ii\alpha S
  =
  -2\ii\alpha C_{\rm zm}.
\]
By multilinearity of the determinant,
\[
  \delta_\alpha\det C_{\rm zm}
  =
  -4\ii\alpha\,\det C_{\rm zm}
\]
for two flavors.  The anomalous phase varies as
\(\delta_\alpha\ee^{\ii\theta}=4\ii\alpha\ee^{\ii\theta}\).  The two
variations cancel, proving
\eqref{eq:instanton-source-determinant-ward-transport}.  This is the
source-determinant form of the local Ward identity proved in
Proposition~\ref{prop:two-flavor-instanton-source-ward-ledger}.
\end{proof}

This proposition is not a new moduli-space statement.  It is the finite
zero-mode algebra which keeps the theta phase, the mass/source determinant,
and the chosen source frame on the same anomalous Ward orbit.  Rotating only
the source determinant changes a charged hard coefficient; shifting only
\(\theta\) changes the sector phase; and applying a fixed pole, OPE, or
susceptibility projection after changing the source frame changes the
observable coordinate.  The Ward identity becomes a physical statement only
after the projection map has also been named and transported in the same
source convention.

\section{A Hard Four-Fermion Benchmark}
\label{sec:instanton-hard-four-fermion-benchmark}

The cleanest test of the preceding distinction is the hard two-flavor
four-fermion channel.  Work in \(SU(3)\) with two light Dirac fundamentals,
choose a \(Q=1\) BPST sector, and insert four amputated external fermion
sources with momenta of size \(Q\).  The center integral first imposes
momentum conservation.  The color orientation integral then projects the
four source slots through a shared Haar tensor.  The fermion zero modes
produce two rank conditions, one for the right-handed overlap matrix and one
for the left-handed overlap matrix.  Only after the conservation, Haar,
zero-mode-rank, and amputation data are fixed does the size integral define
the hard coefficient.

\begin{proposition}[Color-orientation tensor in the hard channel]
\label{prop:instanton-hard-haar-orientation-tensor}
For a charge-one \(SU(N_c)\) instanton, choose an oriented color two-frame
\((u,v)\) describing the embedded \(SU(2)\) color plane, normalized by
\[
  u^\dagger u=v^\dagger v=1,\qquad u^\dagger v=0 .
\]
Let
\[
  W_{ab}=u_av_b-u_bv_a .
\]
With the invariant orientation measure normalized to total mass one, the
two-frame average is
\begin{equation}
  \int \dd\Omega_{N_c}\,
  W_{ab}\,\overline{W_{cd}}
  =
  \frac{2}{N_c(N_c-1)}
  \bigl(
    \delta_{ac}\delta_{bd}
    -
    \delta_{ad}\delta_{bc}
  \bigr).
  \label{eq:instanton-hard-haar-two-frame-tensor}
\end{equation}
Consequently, if \(X_{ab}\) and \(Y_{ab}\) are the two color-pair source
tensors entering a retained hard channel, the color-orientation part of the
amplitude is the antisymmetric projection
\begin{equation}
  \mathcal P_{\rm Haar}(X,Y)
  =
  \frac{2}{N_c(N_c-1)}
  \sum_{a,b}
  X_{ab}\,
  \bigl(\overline{Y_{ab}}-\overline{Y_{ba}}\bigr).
  \label{eq:instanton-hard-haar-source-projection}
\end{equation}
A color-symmetric pair source therefore gives zero in this channel even if
the BPST density, nonzero-mode determinant, zero-mode flavor determinants,
and hard size window are all nonzero.
\end{proposition}

\begin{proof}
The average
\[
  T_{ab,cd}
  =
  \int \dd\Omega_{N_c}\,W_{ab}\overline{W_{cd}}
\]
is invariant under the global color group and is antisymmetric in
\((a,b)\) and in \((c,d)\).  On
\(\wedge^2\mathbb C^{N_c}\) the only invariant tensor with these symmetries
is a scalar multiple of
\(\delta_{ac}\delta_{bd}-\delta_{ad}\delta_{bc}\).  Thus
\[
  T_{ab,cd}
  =
  C_{N_c}
  \bigl(
    \delta_{ac}\delta_{bd}
    -
    \delta_{ad}\delta_{bc}
  \bigr).
\]
Contracting \(a=c\) and \(b=d\) fixes the coefficient.  The left side gives
\[
  \int\dd\Omega_{N_c}\sum_{a,b}|W_{ab}|^2
  =
  2,
\]
because \(u\) and \(v\) are an orthonormal two-frame.  The right side gives
\[
  C_{N_c}(N_c^2-N_c).
\]
Therefore \(C_{N_c}=2/[N_c(N_c-1)]\), which proves
\eqref{eq:instanton-hard-haar-two-frame-tensor}.  Contracting this tensor
with \(X_{ab}\overline{Y_{cd}}\) gives
\eqref{eq:instanton-hard-haar-source-projection}.  If \(Y_{ab}=Y_{ba}\),
the parenthesis in \eqref{eq:instanton-hard-haar-source-projection} vanishes
pointwise.  The vanishing is a color-projection statement, not a statement
about the collective-coordinate density.
\end{proof}

\begin{proposition}[Individual zero-mode slot in the hard channel]
\label{prop:instanton-hard-individual-zero-mode-slot}
The hard four-source channel uses the Fourier transform of individual BPST
fermion zero modes.  In the singular-gauge slot normalization of
Proposition~\ref{prop:instanton-individual-zero-mode-form-factor}, after the
spin-color tensor and the adjacent amputation convention have been separated,
the scalar slot profile is
\begin{equation}
  F_{\rm zm}(t)
  =
  -t\frac{\dd}{\dd t}
  \bigl[I_0(t)K_0(t)-I_1(t)K_1(t)\bigr]
  =
  2t\bigl[I_0(t)K_1(t)-I_1(t)K_0(t)\bigr]
  -2I_1(t)K_1(t),
  \label{eq:instanton-physical-individual-slot-form-factor}
\end{equation}
where \(t=\rho |p|/2\).  It has local limit \(F_{\rm zm}(0)=1\), while
\begin{equation}
  F_{\rm zm}(t)
  =
  \frac{3}{4t^3}+O(t^{-5})
  =
  \frac{6}{(\rho |p|)^3}
  +O((\rho |p|)^{-5}).
  \label{eq:instanton-physical-individual-slot-tail}
\end{equation}
Thus, for four hard slots with \(|p_\ell|=c_\ell Q\) and \(s=Q\rho\),
\begin{equation}
  \prod_{\ell=1}^4 F_{\rm zm}(c_\ell s/2)
  =
  6^4
  \left(\prod_{\ell=1}^4c_\ell^{-3}\right)
  s^{-12}
  +O(s^{-14})
  \label{eq:instanton-physical-four-slot-product-tail}
\end{equation}
when the \(c_\ell\)'s are held fixed away from zero.  A fused bilinear-density
source instead uses \(zK_1(z)\) and has exponential large-\(z\) decay.  The
two source types therefore define different endpoint problems before any
physical projection is taken.
\end{proposition}

\begin{proof}
The first equality in
\eqref{eq:instanton-physical-individual-slot-form-factor} is the
singular-gauge Fourier transform of the normalized fundamental zero mode
after its spin-color tensor has been factored out.  Differentiating
\(\Phi(t)=I_0(t)K_0(t)-I_1(t)K_1(t)\), using
\[
  I_0'=I_1,\qquad
  I_1'=I_0-\frac{I_1}{t},\qquad
  K_0'=-K_1,\qquad
  K_1'=-K_0-\frac{K_1}{t},
\]
gives the second expression in
\eqref{eq:instanton-physical-individual-slot-form-factor}.  The small-\(t\)
expansions \(I_0=1+O(t^2)\), \(I_1=t/2+O(t^3)\), and
\(K_1=t^{-1}+O(t\log t)\) give \(F_{\rm zm}(0)=1\).

For the hard endpoint, use the large-\(t\) products
\[
\begin{aligned}
  I_0K_1
  &=
  \frac1{2t}
  \left(1+\frac1{2t}+\frac{3}{16t^3}+O(t^{-5})\right),\\
  I_1K_0
  &=
  \frac1{2t}
  \left(1-\frac1{2t}-\frac{3}{16t^3}+O(t^{-5})\right),\\
  I_1K_1
  &=
  \frac1{2t}
  \left(1-\frac{3}{8t^2}+O(t^{-4})\right).
\end{aligned}
\]
Then
\[
  2t(I_0K_1-I_1K_0)
  =
  \frac1t+\frac{3}{8t^3}+O(t^{-5}),
  \qquad
  2I_1K_1
  =
  \frac1t-\frac{3}{8t^3}+O(t^{-5}).
\]
The \(t^{-1}\) terms cancel in \(F_{\rm zm}\), leaving
\(3/(4t^3)+O(t^{-5})\).  Substituting \(t=c_\ell s/2\) gives the
single-slot coefficient \(6c_\ell^{-3}s^{-3}\), and multiplying the four
fixed hard slots gives
\eqref{eq:instanton-physical-four-slot-product-tail}.
\end{proof}

Let \(Q_\ell=c_\ell Q\), \(c_\ell>0\), be the four hard source scales, and let
\(F_\ell(c_\ell Q\rho)\) denote the corresponding individual zero-mode slot
form factor, including any bounded source-window factor in the declared
source normalization.  After the center delta, common determinant
constant, Haar contraction, amputation convention, and source-overlap
determinants have been pulled into a channel prefactor
\(\mathcal K_{\rm ch}\), the \(SU(3)\) bosonic zero-mode normalizations still
leave the running collective factor
\[
  \Gamma_{\rm coll}(Q)
  :=
  \left(\frac{8\pi^2}{g_{\rm ht}^2(Q)}\right)^6 .
\]
This factor is not part of the finite determinant constant and cannot be
absorbed into a scale-independent \(\mathcal K_{\rm ch}\).  The stripped
one-instanton hard coefficient has the form
\begin{equation}
  \mathcal A_{4}^{\rm hard}(Q)
  =
  \mathcal K_{\rm ch}\,
  \Gamma_{\rm coll}(Q)\,
  \Lambda_{\rm ht}^{29/3}\,
  Q^{-35/3}\,
  H(c_1,c_2,c_3,c_4)
  +R_{4}^{\rm hard}(Q),
  \label{eq:instanton-hard-four-source-coefficient}
\end{equation}
where
\begin{equation}
  H(c_1,c_2,c_3,c_4)
  =
  \int_0^\infty
  s^{32/3}\prod_{\ell=1}^{4}F_\ell(c_\ell s)\,\dd s .
  \label{eq:instanton-hard-four-source-window-integral}
\end{equation}
The residual \(R_{4}^{\rm hard}\) contains the determinant, source-frame,
amputation, endpoint, sector-isolation, continuation, and physical-projection
remainders from \eqref{eq:instanton-physical-channel-residual-budget}.  The
formula is a benchmark channel, not a universal local vertex: changing the
source wave packets, the orientation contraction, or the physical projection
changes \(\mathcal K_{\rm ch}\), \(H\), or \(R_{4}^{\rm hard}\).

\begin{controlledapproximation}[Amputated 't Hooft four-point assembly]
\label{ca:instanton-thooft-four-point-amputated-assembly}
The hard four-source coefficient is the finite-regulator remnant of an
amputated Green-function calculation.  Before the center integral is
evaluated, a retained \(Q=1\) contribution has the ordered form
\begin{equation}
\begin{aligned}
  \mathcal G_{4,\Lambda}^{I}(p_1,\ldots,p_4)
  &=
  \ee^{\ii\theta}
  \int
  \dd^4a\,\ee^{\ii a\cdot(p_1+\cdots+p_4)}
  \int_{\mathcal W_\rho}
  \dd\rho\int\dd\Omega_{3}\,
  D_{\Lambda}^{I}(\rho,\Omega)                         \\
  &\quad{}\times
  \mathcal H_{\Omega}(\mathcal J_R,\mathcal J_L)\,
  \det B_R(\rho,\Omega;\mathcal J)\,
  \det B_L(\rho,\Omega;\mathcal J)
  \prod_{\ell=1}^{4}
    \Psi_{\ell}^{\rm amp}(\rho p_\ell)                  \\
  &\quad{}\times
  \mathcal R_{\rm nz-src}(\rho,\Omega;\mathcal J)\,
  \mathcal R_{\rm amp}(\rho,\Omega;\mathcal J)\,
  \mathcal R_{\rm proj}(\rho,\Omega;\mathcal J).
\end{aligned}
  \label{eq:instanton-thooft-four-point-amputated-assembly}
\end{equation}
Here \(D_{\Lambda}^{I}\) contains the classical action, collective-coordinate
Jacobian, and zero-mode-deleted determinant in one regulator convention;
\(\mathcal H_{\Omega}\) is the shared color-orientation tensor; \(B_R\) and
\(B_L\) are the right- and left-handed zero-mode source matrices; the
\(\Psi_{\ell}^{\rm amp}\) are the individual amputated zero-mode slots; and
the last three factors record, respectively, the source-dependent nonzero-mode
quotient, external-leg amputation transport, and the chosen physical
projection.  The \(a\)-integral gives
\((2\pi)^4\delta^{(4)}(\sum_\ell p_\ell)\); only after this projection has
been applied does the remaining size/orientation integral define the hard
coefficient in \eqref{eq:instanton-hard-four-source-coefficient}.

The difficult part of the original calculation is that all entries in
\eqref{eq:instanton-thooft-four-point-amputated-assembly} must be normalized
in the same regulator.  The one-loop density fixes \(D_{\Lambda}^{I}\), but
it does not fix the chiral source determinants, the antisymmetric Haar
projection, the amputated zero-mode form factors, or the normal-fluctuation
source quotient.  Conversely, the zero-mode determinant fixes the anomalous
source selection but not the heat-kernel determinant, the finite orientation
normalization, or the endpoint tail.

For a controlled channel write the comparison with the physical four-point
coordinate as
\begin{equation}
\begin{aligned}
  \mathcal A_{4}^{\rm phys}
  -
  \mathcal A_{4,\Lambda}^{\rm amp}
  &=
  E_{\rm dens}
  +E_{\rm slot}
  +E_{\rm Haar}
  +E_{\rm src}
  +E_{\rm nz-src}                                      \\
  &\qquad
  +E_{\rm amp}
  +E_{\rm tail}
  +E_{\rm proj}
  +E_{\rm sec},
\end{aligned}
  \label{eq:instanton-thooft-four-point-residual}
\end{equation}
with absolute bound
\begin{equation}
  \left|
  \mathcal A_{4}^{\rm phys}
  -
  \mathcal A_{4,\Lambda}^{\rm amp}
  \right|
  \le
  B_{\rm dens}+B_{\rm slot}+B_{\rm Haar}+B_{\rm src}
  +B_{\rm nz-src}+B_{\rm amp}+B_{\rm tail}+B_{\rm proj}+B_{\rm sec}.
  \label{eq:instanton-thooft-four-point-bound}
\end{equation}
The terms are not interchangeable.  \(E_{\rm dens}\) controls the density and
determinant normalization; \(E_{\rm slot}\) the individual zero-mode
wavefunctions; \(E_{\rm Haar}\) the color orientation projection;
\(E_{\rm src}\) the chiral source-rank and flavor determinant;
\(E_{\rm nz-src}\) the nonzero-mode source quotient; \(E_{\rm amp}\) external-leg
amputation and source-overlap transport; \(E_{\rm tail}\) the size-window
endpoint; \(E_{\rm proj}\) the pole, OPE, spectral, or inclusive projection;
and \(E_{\rm sec}\) sector isolation from neighboring instanton number.  A
density-only result sets none of these later terms to zero.  A
zero-mode-only result sets neither the determinant nor projection terms to
zero.  The physical four-point claim is the assembled amplitude together with
the residual bound.
\end{controlledapproximation}

\begin{controlledapproximation}[Crossed chiral channel from the 't Hooft kernel]
\label{ca:instanton-thooft-crossed-chiral-channel}
The four-source kernel in
ControlledApproximation~\ref{ca:instanton-thooft-four-point-amputated-assembly}
is still an all-outgoing source coordinate.  A scattering channel is obtained
only after the zero-mode chirality labels have first selected the source
coefficient and the selected external slots have then been crossed with their
LSZ residues.  In the \(Q=1\), \(N_f=2\) convention used above, the supported
all-outgoing source monomial is
\[
  q_{L,u}(p_1)\,q_{L,d}(p_2)\,
  \bar q_{R,u}(p_3)\,\bar q_{R,d}(p_4).
\]
Crossing the two barred slots, with \(p_3=-k_3\) and \(p_4=-k_4\), gives the
chiral physical channel
\begin{equation}
  u_R(k_3)+d_R(k_4)
  \longrightarrow
  u_L(p_1)+d_L(p_2).
  \label{eq:instanton-thooft-crossed-channel-process}
\end{equation}
Thus the physical channel is not obtained by erasing the all-outgoing labels.
It is
\begin{equation}
  \mathcal M_I^{R R\to L L}
  =
  \Xi_{\rm cr}(k_3,k_4;p_1,p_2)\,
  \mathcal A_{I}^{\rm all\text{-}out}
       (p_1,p_2,-k_3,-k_4)
  +R_{\rm cr},
  \label{eq:instanton-thooft-crossed-channel-amplitude}
\end{equation}
where \(\Xi_{\rm cr}\) contains the spinor crossing signs, wave-packet
normalizations, and pole residues in the same source basis used for
amputation.  The amplitude inherits the phase and anomalous source selection
\begin{equation}
\begin{aligned}
  \mathcal M_I^{R R\to L L}
  &=
  \ee^{\ii\theta}K_{R R\to L L}+R_{\rm cr},
  \\
  K_{R R\to L L}
  &=
  \Xi_{\rm cr}\,
  \mathcal C_I\,
  \det B_R\,\det B_L\,
  \mathcal H_\Omega                                      \\
  &\quad{}\times
  \prod_{\ell=1}^{4}\Psi_\ell^{\rm amp}
  \mathcal R_{\rm nz-src}\mathcal R_{\rm amp}\mathcal R_{\rm proj},
  \label{eq:instanton-thooft-crossed-channel-kernel}
\end{aligned}
\end{equation}
with the same density and size-window convention as
\eqref{eq:instanton-thooft-four-point-amputated-assembly}.  The anti-instanton
sector gives the conjugate chirality channel
\[
  u_L+d_L\longrightarrow u_R+d_R,
  \qquad
  \mathcal M_{\bar I}^{L L\to R R}
  =
  \ee^{-\ii\theta}K_{L L\to R R}^{\bar I}
  +R_{\rm cr}^{\bar I}.
\]
For massless external quarks these are different chiral channels.  A
perturbative amplitude that preserves chirality is not a reference amplitude in
\eqref{eq:instanton-thooft-crossed-channel-process}; a linear
\(I+\bar I\) sum is therefore not a positive rate and not an inclusive cut.
The selected channel rate has the quadratic neutral pairing
\begin{equation}
  |\mathcal M_I^{R R\to L L}|^2
  =
  |K_{R R\to L L}|^2
  +O(KR_{\rm cr})+O(R_{\rm cr}^2),
  \label{eq:instanton-thooft-crossed-channel-rate}
\end{equation}
so the one-instanton \(\theta\)-phase cancels in the positive quadratic
projection.  A \(\theta\)-linear observable requires a declared chirality
breaking reference channel, mass insertion, or CP-paired comparison with its
own projection and residual budget.

Consequently the hard four-point calculation has three separate decisions:
which all-outgoing anomalous source monomial is differentiated, which slots
are crossed into the physical in-state, and which quadratic or interference
observable is measured.  Crossing before the zero-mode chirality selection
mixes the \(Q=1\) and \(Q=-1\) determinants.  Summing the instanton and
anti-instanton amplitudes before specifying the measurement mixes conjugate
chiral processes.  Reading a chirality-balanced external process as a hard
't Hooft amplitude sets the source determinant in the wrong coordinate.
\end{controlledapproximation}

\begin{controlledapproximation}[Mass-assisted interference channel]
\label{ca:instanton-mass-assisted-interference-channel}
The mixed terms in
\eqref{eq:two-flavor-mass-source-determinant-polynomial} are the first place
where the theta-charged instanton amplitude can meet an ordinary
chirality-breaking reference in the same external channel.  Let
\[
  m_f=|m_f|\ee^{\ii\phi_f}
\]
be the complex mass coordinate in the same zero-mode block as
Proposition~\ref{prop:two-flavor-source-mass-determinant-coordinate}.  The
\(u\)-flavor two-source coordinate of the \(Q=1\) sector is the coefficient of
\(B_{uu}\):
\begin{equation}
  \mathcal M_{I,u}^{R\to L}
  =
  \ee^{\ii\theta}\,
  m_d\,K_{I,u}\,\Xi_u
  +R_{I,u}.
  \label{eq:instanton-mass-assisted-u-channel}
\end{equation}
Here \(m_d\) saturates the \(d\)-flavor zero modes, while \(B_{uu}\), after
crossing, supplies the external \(u_R\to u_L\) state.  The analogous
\(d\)-flavor channel has \(m_uB_{dd}\).  Neither channel is the massless
four-source coefficient \(\det B\), and neither is the vacuum activity
\(m_um_d\).

Suppose a perturbative or long-distance chirality-breaking reference amplitude
has been projected into the same \(u_R\to u_L\) basis:
\begin{equation}
  \mathcal M_{{\rm ref},u}^{R\to L}
  =
  \bar m_u\,K_{{\rm ref},u}
  +R_{{\rm ref},u}.
  \label{eq:instanton-mass-assisted-reference}
\end{equation}
The complex conjugate on \(m_u\) is a convention statement for this
\(R\to L\) external orientation; reversing the external chirality reverses all
conjugations.  In this convention the interference part of a same-basis
measurement is
\begin{equation}
\begin{aligned}
  \mathcal I_u
  &:=
  2\Re\left[
    \overline{\mathcal M_{{\rm ref},u}^{R\to L}}\,
    W_u\,
    \mathcal M_{I,u}^{R\to L}
  \right]                                                \\
  &=
  2|m_um_d|\,
  |K_{{\rm ref},u}W_uK_{I,u}\Xi_u|\,
  \cos\!\left(\theta+\phi_u+\phi_d+\delta_u\right)
  +R_{{\rm int},u},
  \label{eq:instanton-mass-assisted-interference}
\end{aligned}
\end{equation}
where \(\delta_u=\arg(K_{I,u}\Xi_u)-\arg(K_{{\rm ref},u}W_u)\) is a physical
channel phase fixed by the same projection and infrared convention.  Thus the
theta-linear observable is not the massless instanton amplitude by itself.  It
is a same-channel interference between a mass-assisted instanton coefficient
and a chirality-breaking reference, and it carries the invariant phase
\(\theta+\arg m_u+\arg m_d\) up to the channel phase \(\delta_u\).

The residual is bounded in the same measurement norm:
\begin{equation}
  |R_{{\rm int},u}|
  \le
  2|W_u|\,
  \left(
    |\mathcal M_{{\rm ref},u}^{0}|B_{I,u}
    +|\mathcal M_{I,u}^{0}|B_{{\rm ref},u}
    +B_{{\rm ref},u}B_{I,u}
  \right)
  +B_{W,u}.
  \label{eq:instanton-mass-assisted-interference-bound}
\end{equation}
The superscript \(0\) denotes the retained leading amplitudes in
\eqref{eq:instanton-mass-assisted-u-channel} and
\eqref{eq:instanton-mass-assisted-reference}.  The term \(B_{W,u}\) records
measurement-basis and infrared-completeness errors.  If the reference
amplitude preserves chirality, or if it is expressed in a different source
basis, then the interference matrix element is not defined.  If one uses
\(m_uK_{{\rm ref},u}\) instead of \(\bar m_uK_{{\rm ref},u}\) in this
orientation, the phase becomes \(\theta-\phi_u+\phi_d\), which is not the
two-flavor axial invariant.  If one omits the complementary mass \(m_d\), the
zero-mode Berezin integral has not been saturated.
\end{controlledapproximation}

\begin{proposition}[\(SU(3)\), \(N_f=2\) hard-source power and slow tail]
\label{prop:su3-nf2-hard-source-power-slow-tail}
In the preceding hard channel, \(b_0=29/3\).  The four differentiated
zero-mode slots contribute the local source power \(\rho^6\), so the size
integrand before hard form factors is
\begin{equation}
  \rho^{\,b_0+6-5}\,\dd\rho
  =
  \rho^{32/3}\,\dd\rho .
  \label{eq:instanton-hard-four-source-rho-power}
\end{equation}
After \(s=Q\rho\), the hard coefficient carries the scale power
\begin{equation}
  \Gamma_{\rm coll}(Q)\,
  \Lambda_{\rm ht}^{29/3}Q^{-35/3},
  \qquad
  29/3-35/3=-2 ,
  \label{eq:instanton-hard-four-source-scale-power}
\end{equation}
where the displayed \(Q^{-35/3}\) is the power-counting part and
\(\Gamma_{\rm coll}(Q)\) is a logarithmically running collective-coordinate
Jacobian.  The mass dimension is that of a four-fermion coefficient.  If all
four individual zero-mode form factors are the singular-gauge hard slots in
Proposition~\ref{prop:instanton-hard-individual-zero-mode-slot}, so that
\[
  F_\ell(c_\ell s)=6c_\ell^{-3}s^{-3}+O(s^{-5}),
\]
then the large-\(s\) endpoint of \eqref{eq:instanton-hard-four-source-window-integral}
has leading tail
\begin{equation}
  \int_R^\infty
  s^{32/3}\prod_{\ell=1}^{4}F_\ell(c_\ell s)\,\dd s
  =
  3\,6^4\!\left(\prod_{\ell=1}^{4}c_\ell^{-3}\right)R^{-1/3}
  +O(R^{-7/3}).
  \label{eq:instanton-hard-four-source-slow-tail}
\end{equation}
If one of the four slots is not hard, this power test no longer controls the
large-size endpoint by the same source profile.
\end{proposition}

\begin{proof}
For \(SU(3)\) with two Dirac fundamentals,
\[
  b_0=\frac{11}{3}3-\frac{2}{3}2=\frac{29}{3}.
\]
The universal one-loop size density contributes \(\rho^{b_0-5}\dd\rho\), and
the four differentiated zero-mode slots contribute \(\rho^6\).  This gives
\eqref{eq:instanton-hard-four-source-rho-power}.  The change of variables
\(s=Q\rho\) gives
\[
  \dd\rho\,\rho^{32/3}
  =
  Q^{-35/3}s^{32/3}\,\dd s .
\]
At the hard subtraction point \(\mu=Q\), the one-loop density prefactor
\(\mu^{b_0}\) contributes \(Q^{29/3}\), while the bosonic zero-mode
normalizations contribute
\(\Gamma_{\rm coll}(Q)=(8\pi^2/g_{\rm ht}^2(Q))^6\).  Therefore before
rewriting the classical exponential the coefficient scales as
\(\exp[-8\pi^2/g_{\rm ht}^2(Q)]Q^{-2}\).  The one-loop RG-invariant trade
\[
  \exp[-8\pi^2/g_{\rm ht}^2(Q)]
  =
  \Lambda_{\rm ht}^{29/3}Q^{-29/3}
\]
then gives the full leading scale dependence
\(\Gamma_{\rm coll}(Q)\Lambda_{\rm ht}^{29/3}Q^{-35/3}\).  The logarithmic
factor is dimensionless; the total mass dimension is \(29/3-35/3=-2\), as
required for a coefficient multiplying a dimension-six four-fermion source.
At one-loop order
\[
  -\frac{\dd}{\dd\log Q}
  \log\!\left[\Gamma_{\rm coll}(Q)Q^{-35/3}\right]
  =
  \frac{35}{3}
  -6\,\frac{\dd}{\dd\log Q}
  \log\!\left(\frac{8\pi^2}{g_{\rm ht}^2(Q)}\right),
\]
so the pure \(35/3\) logarithmic slope is a power-counting statement modulo
the collective-mode logarithm.

For the endpoint, multiply the four large-\(s\) tails.  The product contributes
\[
  6^4\left(\prod_\ell c_\ell^{-3}\right)s^{-12}+O(s^{-14}).
\]
Multiplication by \(s^{32/3}\) gives
\[
  6^4\left(\prod_\ell c_\ell^{-3}\right)s^{-4/3}+O(s^{-10/3}).
\]
Integrating from \(R\) to infinity gives
\eqref{eq:instanton-hard-four-source-slow-tail}.  With only three hard slots
the leading endpoint power is \(s^{32/3-9}=s^{5/3}\), which is not integrable
at infinity.
\end{proof}

\begin{controlledapproximation}[Tail-subtracted hard-window evaluation]
\label{ca:instanton-hard-window-tail-subtraction}
The slow \(R^{-1/3}\) endpoint in
\eqref{eq:instanton-hard-four-source-slow-tail} is not a reason to abandon
the hard benchmark; it is a reason to evaluate the benchmark as a
tail-subtracted amplitude.  For the same singular-gauge differentiated slots,
the next term in the large-\(s\) expansion is
\begin{equation}
  F_\ell(c_\ell s)
  =
  6c_\ell^{-3}s^{-3}
  +
  45c_\ell^{-5}s^{-5}
  +O(s^{-7}).
  \label{eq:instanton-hard-slot-two-term-tail}
\end{equation}
Set
\[
  A_0=6^4\prod_{\ell=1}^4c_\ell^{-3},
  \qquad
  A_1
  =
  45\,6^3
  \sum_{m=1}^{4}
  c_m^{-5}\prod_{\ell\ne m}c_\ell^{-3}.
\]
Then
\begin{equation}
  s^{32/3}\prod_{\ell=1}^4F_\ell(c_\ell s)
  =
  A_0s^{-4/3}+A_1s^{-10/3}+O(s^{-16/3}),
  \label{eq:instanton-hard-window-two-term-integrand-tail}
\end{equation}
and the physical hard-window integral can be written as
\begin{equation}
\begin{aligned}
  H(c_1,\ldots,c_4)
  &=
  \int_0^R
    s^{32/3}\prod_{\ell=1}^{4}F_\ell(c_\ell s)\,\dd s       \\
  &\quad
  +3A_0R^{-1/3}
  +\frac37A_1R^{-7/3}
  +E_{\rm tail}(R),
\end{aligned}
  \label{eq:instanton-hard-window-tail-subtracted-integral}
\end{equation}
with \(E_{\rm tail}(R)=O(R^{-13/3})\) in a window where the
\(O(s^{-7})\) remainder in \eqref{eq:instanton-hard-slot-two-term-tail} is
uniform for the chosen ratios \(c_\ell\).  Thus the hard amplitude is not just
a formal \(\rho\)-integral: the short numerical core, the leading endpoint
tail, the subleading endpoint tail, and the residual estimate are four
separate pieces of the physical coefficient.

This calculation also shows what must not be done.  Dropping the
\((3/7)A_1R^{-7/3}\) term is not a change of convention; it changes the
retained hard-window coefficient at the same source normalization.  Replacing
the differentiated zero-mode slots by a fused \(zK_1(z)\) density changes the
endpoint class rather than improving the same integral.  Finally, no value of
\(H\) by itself is an on-shell hadronic amplitude; it still has to be inserted
into the assembled channel and then into the Wilsonian, pole, spectral, or
inclusive projection described below.
\end{controlledapproximation}

\begin{controlledapproximation}[Screened hard-size majorant window]
\label{ca:instanton-hard-screened-retained-window}
The tail subtraction in
\eqref{eq:instanton-hard-window-tail-subtracted-integral} controls the power
endpoint of the differentiated zero-mode slots.  It does not, by itself, show
that the retained instanton sizes dominate a physical amplitude.  Such a
statement needs a channel-specific long-size mechanism from the same regulated
observable: for example a Higgs scale, a thermal/medium scale, a finite-volume
gap, a spectral-bin source envelope, or a Wilsonian matrix element estimate.

Fix the selected source windows \(U,V\), the color projection, the amputation
map, and the determinant scheme before the final physical projection.  Let
\(\mathcal K^{\rm pre}_{\mathcal C,\Lambda}(\rho;U,V)\) denote the resulting
pre-projection radial kernel, and let \(P_{\rm phys}^{\mathcal C}\) be the
linear projection to the declared physical coordinate.  Assume the source
windows have dimensionless norm \(M_{\rm src}(\mathcal C)\), that
\(\|P_{\rm phys}^{\mathcal C}\|=M_{\rm phys}(\mathcal C)\), and that the
pre-projection kernel has the absolute majorant
\begin{equation}
  |\mathcal K^{\rm pre}_{\mathcal C,\Lambda}(\rho;U,V)|
  \le
  K_{\mathcal C}\,
  M_{\rm src}(\mathcal C)\,
  \rho^{A_{\mathcal C}-1}
  \exp[-d_{\mathcal C}Q\rho-m_{\rm scr}^2\rho^2],
  \qquad
  A_{\mathcal C}>1,\quad d_{\mathcal C}>0 .
  \label{eq:instanton-hard-screened-channel-majorant}
\end{equation}
Here \(A_{\mathcal C}\) is the logarithmic size-shell power before the
exponential factors.  For the raw \(SU(3)\), \(N_f=2\) hard four-source density
before the large-\(s\) zero-mode tails are inserted,
\(\rho^{32/3}\dd\rho=\rho^{35/3}\dd\log\rho\), so
\(A_{\mathcal C}=35/3\).  After the four individual singular-gauge tails have
been inserted, the endpoint shell has power \(-1/3\), and the right tool is the
tail-subtracted evaluation above, not a positive saddle estimate.  The
coefficient \(d_{\mathcal C}Q\) is a proven hard-source envelope for the
selected source/projection window; it is not supplied by the BPST moduli
measure alone.  Likewise \(m_{\rm scr}\) is a physical screening scale in the
same regulated problem, not a hand cutoff placed on the size integral.

The logarithmic shell of the majorant is
\[
  \rho^{A_{\mathcal C}}
  \exp[-d_{\mathcal C}Q\rho-m_{\rm scr}^2\rho^2]\,\dd\log\rho .
\]
Its stationary point is fixed by the combined equation
\begin{equation}
  2m_{\rm scr}^2\rho_*^2+d_{\mathcal C}Q\rho_*-A_{\mathcal C}=0,
  \qquad
  \rho_*=
  \frac{2A_{\mathcal C}}
       {d_{\mathcal C}Q+
        \sqrt{d_{\mathcal C}^{\,2}Q^2+8A_{\mathcal C}m_{\rm scr}^2}} .
  \label{eq:instanton-hard-screened-log-shell}
\end{equation}
This \(\rho_*\) is the stationary point of the majorizing envelope.  It is not,
without more input, the peak of the physical kernel or the dominant instanton
size in the amplitude.  A signed or complex source amplitude can obey
\eqref{eq:instanton-hard-screened-channel-majorant} while vanishing by
projection, oscillating through the shell, or concentrating near a boundary.
To promote the envelope shell to a dominance statement one must add a
channel-specific lower/comparability estimate, for example a positive spectral
or rate density satisfying a lower bound on a window
\(\mathcal W=[\rho_-,\rho_+]\), together with a quantified tail fraction outside
\(\mathcal W\) and a noncancellation condition after \(P_{\rm phys}^{\mathcal C}\).
The hard-only shell \(A_{\mathcal C}/(d_{\mathcal C}Q)\) and the
screened-only shell \(\sqrt{A_{\mathcal C}/(2m_{\rm scr}^2)}\) are limiting
checks on this equation, not substitutes for it.  Forgetting the passage from
\(\dd\rho\) to \(\dd\log\rho\), replacing \(A_{\mathcal C}=35/3\) by
\(32/3\) in the hard four-source channel, or dropping either the hard envelope
or the physical screening scale moves the shell and changes the amplitude
envelope window.

The long-size lower endpoint and the semiclassical domain are separate gates.
With \(\rho_0=R/Q\), an interior retained-window reading requires
\[
  \rho_0\le \rho_-<\rho_*<\rho_+,
  \qquad
  \Lambda_{\rm QCD}\rho_+\ll1
\]
or the corresponding weakly coupled Higgs/thermal regime that supplied
\(m_{\rm scr}\).  If \(\rho_0\ge\rho_*\), the envelope is already past its
logarithmic shell in the integration domain, and the estimate is controlled by
the boundary rather than an interior saddle.  In that case, whenever
\[
  d_{\mathcal C}Q+2m_{\rm scr}^2\rho_0
  -\frac{A_{\mathcal C}-1}{\rho_0}>0,
\]
convexity of
\(
d_{\mathcal C}Q\rho+m_{\rm scr}^2\rho^2-(A_{\mathcal C}-1)\log\rho
\)
gives the boundary tail bound
\[
  \int_{\rho_0}^{\infty}
  \rho^{A_{\mathcal C}-1}
  e^{-d_{\mathcal C}Q\rho-m_{\rm scr}^2\rho^2}\,\dd\rho
  \le
  \frac{
  \rho_0^{A_{\mathcal C}-1}
  e^{-d_{\mathcal C}Q\rho_0-m_{\rm scr}^2\rho_0^2}
  }{
  d_{\mathcal C}Q+2m_{\rm scr}^2\rho_0
  -(A_{\mathcal C}-1)/\rho_0
  } .
\]

Consequently the long-size contribution is controlled only after the source
and projection norms are kept with the size estimate:
\begin{equation}
\begin{aligned}
  |\mathcal A^I_{\mathcal C,>}(Q,R)|
  &\le
  K_{\mathcal C}\,
  M_{\rm src}(\mathcal C)\,
  M_{\rm phys}(\mathcal C)
  \int_{R/Q}^{\infty}
    \rho^{A_{\mathcal C}-1}
    e^{-d_{\mathcal C}Q\rho-m_{\rm scr}^2\rho^2}
    \,\dd\rho                                      \\
  &\quad
  +E_{\rm det}^{>}+E_{\rm nm}^{>}+E_{\rm tail}^{>}+E_{\rm phys}^{>} .
\end{aligned}
  \label{eq:instanton-hard-screened-window-bound}
\end{equation}
This is the amplitude-level use of the screened majorant.  It asks which
long-size region remains compatible with the hard source envelope, screening
scale, weak-coupling domain, source norms, projection norm, and residual
telescope.  It is not a new moduli-space volume formula, not a proof that the
actual amplitude is concentrated at \(\rho_*\), and not a phenomenological
instanton regime by itself.
A 't Hooft-style four-point calculation must still exhibit the determinant,
nonzero-mode source response, zero-mode slots, amputation, projection, and
residuals in the common regulator before the screened window can be read as a
physical amplitude contribution.
\end{controlledapproximation}

\begin{controlledapproximation}[Hard benchmark channel comparison]
\label{ca:instanton-hard-benchmark-gate-ledger}
A controlled comparison of two hard four-source instanton coefficients at
scales \(Q_1,Q_2\) may use the power in
\eqref{eq:instanton-hard-four-source-scale-power} only after the channel data are
held fixed:
\[
  \Delta_{\rm cen}\ne0,\qquad
  \mathcal P_{\rm Haar}\ne0,\qquad
  \det B_R\ne0,\qquad
  \det B_L\ne0,
\]
with the same amputation convention, source-frame normalization, determinant
scheme, sector projection, and dimensionless source-window shape.  If the
retained coefficients have the form
\[
  \mathcal A_j
  =
  \mathcal A_j^{(0)}(1+\varepsilon_j),
  \qquad
  |\varepsilon_j|\le e_j<1,
  \qquad j=1,2,
\]
and the leading channel data are the same at the two scales, then
\begin{equation}
  \frac{\mathcal A_2}{\mathcal A_1}
  =
  \frac{\Gamma_{\rm coll}(Q_2)}{\Gamma_{\rm coll}(Q_1)}
  \left(\frac{Q_2}{Q_1}\right)^{-35/3}
  \frac{1+\varepsilon_2}{1+\varepsilon_1},
  \label{eq:instanton-hard-four-source-ratio}
\end{equation}
with residual multiplier bounded by
\begin{equation}
  \left|
  \frac{1+\varepsilon_2}{1+\varepsilon_1}-1
  \right|
  \le
  \frac{e_1+e_2}{1-e_1}.
  \label{eq:instanton-hard-four-source-ratio-bound}
\end{equation}
Thus the determinant constant and \(\Lambda_{\rm ht}\) cancel in a same-theory
hard-scale ratio, while the running collective factor, the dimensionless
source-window, and the rank-conditioning data do not cancel unless they have
actually been transported.  An off-shell center projection, a rank-one
zero-mode overlap, an omitted Haar projection, an omitted collective-mode
Jacobian, or unamputated external residues change the amplitude coordinate
before the power law \((Q_2/Q_1)^{-35/3}\) can be tested modulo logarithms.
\end{controlledapproximation}

\section{The Hard Kernel as a Wilsonian OPE Input}
\label{sec:instanton-hard-wilsonian-ope-datum}

The hard coefficient in
\eqref{eq:instanton-hard-four-source-coefficient} is still a source
coefficient.  To use it as a local QFT input, one must decide which part of
the size integral is short-distance and which part is left for the matrix
element or the long-distance remainder.  This is the channel-level version of
the Wilsonian split in
Controlled Approximation~\ref{ca:short-instanton-ope-coefficient}.

Fix a dimensionless size split \(R\) in the hard variable \(s=Q\rho\), and
write
\[
  \rho_I=\frac{R}{Q},
  \qquad
  \mu_I=\frac{Q}{R}.
\]
Let \([O_4]_\mu\) be the renormalized local four-fermion operator obtained
from the same spin, flavor, and color-orientation projection as the hard
source channel.  The short-instanton coefficient is
\begin{equation}
  C_{4}^{<}(Q,R;\mu)
  =
  \ee^{\ii\theta}
  \mathcal K_{\rm ch}\,
  \Gamma_{\rm coll}(Q)\,
  \Lambda_{\rm ht}^{29/3}\,
  Q^{-35/3}
  Z_{4}(R,\mu)
  \int_0^{R}
    s^{32/3}\prod_{\ell=1}^{4}F_\ell(c_\ell s)\,\dd s ,
  \label{eq:instanton-hard-short-ope-coefficient}
\end{equation}
where \(Z_4(R,\mu)\) is the finite matching matrix element from the source
projection to the chosen renormalized operator coordinate.  A physical
hadronic or inclusive observable then has the form
\begin{equation}
  \mathcal A_{\rm phys}^{I}
  =
  C_{4}^{<}(Q,R;\mu)\,
  \langle f|[O_4]_\mu|i\rangle_{\rm phys}
  +
  \mathcal A_{4}^{>}(Q,R)
  +
  R_{\rm match}+R_{\rm IR}+R_{\rm proj}+R_{\rm sec}.
  \label{eq:instanton-hard-ope-physical-assembly}
\end{equation}
The first term is local only in the Wilsonian sense.  The second term is the
same instanton source channel with \(\rho>\rho_I\), including whatever
screening, confinement, finite-volume, or source-window mechanism controls
that region.  The residuals record, respectively, operator matching, infrared
completion, final physical projection, and sector isolation.  Thus
\(C_4^{<}\) is not the physical amplitude unless the matrix element and the
long-size/remainder terms have been supplied.

The factorization-scale flow is a boundary flux of the same size integral:
\begin{equation}
  \partial_{\log\mu_I}C_4^{<}
  =
  -\ee^{\ii\theta}
  \mathcal K_{\rm ch}
  \Gamma_{\rm coll}(Q)
  \Lambda_{\rm ht}^{29/3}
  Q^{-35/3}
  Z_{4}(R,\mu)\,
  R^{35/3}\prod_{\ell=1}^{4}F_\ell(c_\ell R)
  +
  C_4^{<}\,\partial_{\log\mu_I}\log Z_4 .
  \label{eq:instanton-hard-short-ope-boundary-flow}
\end{equation}
The first term is the explicit movement of the shell
\(\rho=R/Q\) between coefficient and remainder; the second is ordinary
operator-coordinate transport.  Changing the local operator basis changes
\(Z_4\), \(C_4^{<}\), and the matrix element inversely, not the completed
observable in \eqref{eq:instanton-hard-ope-physical-assembly}.

\begin{controlledapproximation}[Wilsonian matching covariance for the
instanton vertex]
\label{ca:instanton-wilsonian-matching-covariance}
The short coefficient in
\eqref{eq:instanton-hard-short-ope-coefficient} is a coordinate on a local
operator space.  Let \(O_a(\mu)\) be the retained local operators with the
spin, flavor, and color projection of the hard instanton source, let
\(C_a^{<}(Q,R;\mu)\) be the corresponding row vector of short-size
coefficients, and let
\[
  M_a(\mu)=\langle f|O_a(\mu)|i\rangle_{\rm phys}.
\]
The Wilsonian contribution to the physical observable is
\begin{equation}
  \mathcal A_{\rm loc}^{I}
  =
  C_a^{<}M_a .
  \label{eq:instanton-wilsonian-local-pairing}
\end{equation}
A finite change of renormalized operator coordinates
\[
  O'_a=S_{ab}O_b,
  \qquad S\in GL(r)
\]
acts as
\begin{equation}
  M'=SM,\qquad C'^{<}=C^{<}S^{-1},
  \qquad
  C'^{<}_{a}M'_a=C^{<}_{a}M_a .
  \label{eq:instanton-wilsonian-scheme-covariance}
\end{equation}
Thus the instanton vertex coefficient by itself is scheme dependent.  The
observable coordinate is the paired coefficient and matrix element, plus the
long-size remainder and the bridge residuals in
\eqref{eq:instanton-hard-ope-physical-assembly}.

The same covariance controls the moving size boundary.  Suppose
\(\Phi_a(R)\) is the boundary-flux row vector obtained by evaluating the
hard instanton kernel on the shell \(s=R\), and let \(\gamma_{ab}\) be the
anomalous-dimension matrix in the same operator basis.  A closed retained
split has the finite-dimensional flow
\begin{equation}
  \partial_{\log\mu_I}C^{<}
  =
  -\Phi+C^{<}\gamma,
  \qquad
  \partial_{\log\mu_I}M
  =
  -\gamma M,
  \qquad
  \partial_{\log\mu_I}\mathcal A_{>}^{I}
  =
  \Phi_aM_a .
  \label{eq:instanton-wilsonian-covariant-boundary-flow}
\end{equation}
Consequently
\[
  \partial_{\log\mu_I}
  \bigl(C_a^{<}M_a+\mathcal A_{>}^{I}\bigr)
  =
  (-\Phi+C^{<}\gamma)M+C^{<}(-\gamma M)+\Phi M
  =
  0
\]
before residuals.  The boundary shell removed from the local coefficient is
the shell added to the long-size remainder, while the anomalous-dimension
terms cancel between coefficient and matrix element.

In a truncated calculation write the three flow equations with residuals
\(\delta C\), \(\delta M\), and \(\delta_>\).  With the vector \(1\)-norm for
rows and the vector \(\infty\)-norm for columns,
\begin{equation}
  \left|
  \partial_{\log\mu_I}\mathcal A_{\rm phys}^{I}
  \right|
  \le
  \|\delta C\|_1\|M\|_{\infty}
  +
  \|C^{<}\|_1\|\delta M\|_{\infty}
  +
  |\delta_>| .
  \label{eq:instanton-wilsonian-covariance-residual-bound}
\end{equation}
This is the finite matching version of the instanton-amplitude lesson.  A
fixed short vertex under a moving size boundary, a diagonal rescaling used
when the operators actually mix, or a short coefficient quoted without the
renormalized matrix element and long-size shell gives a different QFT
coordinate from the physical observable.
\end{controlledapproximation}

For the individual hard zero-mode slots of
Proposition~\ref{prop:instanton-hard-individual-zero-mode-slot}, the omitted
tail has the leading form
\begin{equation}
  \mathcal A_{4}^{>}(Q,R)
  =
  \ee^{\ii\theta}
  \mathcal K_{\rm ch}
  \Gamma_{\rm coll}(Q)
  \Lambda_{\rm ht}^{29/3}
  Q^{-35/3}
  \left[
    3\,6^4
    \left(\prod_{\ell=1}^{4}c_\ell^{-3}\right)R^{-1/3}
    +O(R^{-7/3})
  \right]
  +R_{>}(Q,R).
  \label{eq:instanton-hard-ope-long-tail}
\end{equation}
The small power \(R^{-1/3}\) is the warning sign.  A hard source suppresses the
large-instanton endpoint, but it does so slowly for individual zero-mode
slots.  A reliable local 't Hooft-operator statement must therefore choose
\(R\), prove that the short coefficient is in the weak-size window, and keep
the long-size tail or show that the physical observable suppresses it.  Using
the full hard source coefficient as a local operator coefficient, or using the
short coefficient as a physical amplitude, skips this Wilsonian boundary
calculation.

\section{Normal Fluctuations Are Source Data}
\label{sec:instanton-normal-fluctuation-source-data}

The one-loop determinant normalizes the Gaussian integral over nonzero modes.
A selected source amplitude also asks how its insertion changes under those
same normal fluctuations.  This is one of the places where the original
instanton amplitude calculation is harder than a moduli-space calculation:
the determinant of the Hessian and the fluctuation average of the source
insertion are coupled through the same measure but are not the same number.

\begin{controlledapproximation}[Primed determinant response to a source perturbation]
\label{ca:instanton-primed-determinant-source-response}
Fix the finite regulator, the background gauge slice, and the subtraction
scheme.  Let \(K_z\) be the bosonic Hessian on the horizontal normal space
after collective and gauge zero modes have been removed, and let
\(G_z=K_z^{-1}\) on that primed space.  Suppose a source or external
background changes the retained quadratic normal action by the exactly linear
operator family
\[
  K_z^{\rm lin}(\epsilon)=K_z+\epsilon R_z
\]
after the \(O(\epsilon^2)\) local counterterm and any higher source dependence
of the Hessian have been moved to separately named counterterm or interaction
coordinates.  The bosonic zero-mode-deleted determinant ratio in this linear
determinant coordinate is
\[
  \mathcal D_B(\epsilon)
  =
  \left[
    \frac{\det{}'K_z^{\rm lin}(\epsilon)}{\det{}'K_z}
  \right]^{-1/2}
  =
  \det{}'(1+\epsilon G_zR_z)^{-1/2}.
\]
Hence
\begin{equation}
  \log \mathcal D_B(\epsilon)
  =
  -\frac{\epsilon}{2}\operatorname{Tr}'(G_zR_z)
  +
  \frac{\epsilon^2}{4}
  \operatorname{Tr}'(G_zR_zG_zR_z)
  +R_{B,{\rm lin}}^{(3)}(\epsilon).
  \label{eq:instanton-primed-boson-determinant-response}
\end{equation}
For a Dirac nonzero-mode block \(D_z(\epsilon)=D_z+\epsilon S_z\), with
\(G_z^{F}=D_z^{-1}\) on the zero-mode-deleted fermion space,
\begin{equation}
  \log \mathcal D_F(\epsilon)
  =
  \epsilon\operatorname{Tr}'(G_z^{F}S_z)
  -
  \frac{\epsilon^2}{2}
  \operatorname{Tr}'(G_z^{F}S_zG_z^{F}S_z)
  +R_{F,{\rm lin}}^{(3)}(\epsilon).
  \label{eq:instanton-primed-fermion-determinant-response}
\end{equation}
The finite counterterm polynomial must be added in the same regulator
coordinate; it is not a license to drop the primed traces.  If
\(\|X_B\|<1\) and \(\|X_F\|<1\), where
\(X_B=\epsilon G_zR_z\) and \(X_F=\epsilon G_z^{F}S_z\), then in the finite
retained normal space the omitted logarithmic terms of the exactly linear
operator families obey
\[
  |R_{B,{\rm lin}}^{(3)}|
  \le
  \frac{\|X_B\|_1\|X_B\|^2}{6(1-\|X_B\|)},
  \qquad
  |R_{F,{\rm lin}}^{(3)}|
  \le
  \frac{\|X_F\|_1\|X_F\|^2}{3(1-\|X_F\|)} .
\]
If the actual regulated Hessian has a retained nonlinear part, for example
\[
  K_z^{\rm full}(\epsilon)
  =
  K_z+\epsilon R_z+\epsilon^3T_z(\epsilon),
  \qquad
  Y_B=\epsilon^3G_zT_z(\epsilon),
\]
then \eqref{eq:instanton-primed-boson-determinant-response} is still the
linearized determinant response, but the full determinant has the additional
operator-family remainder
\[
  R_{B,{\rm op}}(\epsilon)
  =
  -\frac12
  \operatorname{Tr}'\log
  \left[
    1+(1+X_B)^{-1}Y_B
  \right].
\]
When
\(\|X_B\|<1\) and \(\|Y_B\|/(1-\|X_B\|)<1\),
\[
  |R_{B,{\rm op}}|
  \le
  \frac12
  \frac{\|Y_B\|_1/(1-\|X_B\|)}
       {1-\|Y_B\|/(1-\|X_B\|)} .
\]
The Dirac block has the same form without the factor \(1/2\) and with
\(X_F,Y_F\) in the zero-mode-deleted fermion space.  Thus an \(O(\epsilon^3)\)
operator perturbation is not controlled by the linear log-series bound; it is
a separate source-response remainder.
Thus the source response of the determinant is a primed resolvent
calculation.  It is not fixed by the \(J=0\) determinant constant, by the
dimension of the moduli space, or by an unprimed trace with an arbitrary
regulator placed on the zero modes.  The same prime, gauge slice, and
counterterm convention must be used before this determinant response can be
multiplied by the zero-mode source determinant or projected to a physical
amplitude.
\end{controlledapproximation}

\begin{controlledapproximation}[Nonzero-mode source quotient]
\label{ca:instanton-nonzero-mode-source-quotient}
Fix a retained instanton chart point \(z\), project out collective and gauge
zero modes, and let \(\eta\in\mathcal N_z\) denote the normal fluctuation
coordinate.  Let \(d\gamma_z(\eta)\) be the normalized Gaussian measure
defined by the zero-mode-deleted Hessian, and let
\(V_z(\eta)\) be the cubic-and-higher action together with finite counterterm
remainders in the same normal coordinates.  Write the selected source
insertion, after the explicit zero-mode coefficient has been separated, as
\[
  \mathcal Z_z(\eta;\mathcal J)
  =
  \mathcal Z_z(0;\mathcal J)
  \bigl(1+U_z(\eta;\mathcal J)\bigr),
  \qquad U_z(0;\mathcal J)=0 .
\]
On a chart where the leading zero-mode/source coefficient is nonzero, the
fluctuation factor multiplying the hard channel is
\begin{equation}
  \mathcal R_{\rm fluc}(z;\mathcal J)
  =
  \frac{
    \left\langle
      \bigl(1+U_z(\eta;\mathcal J)\bigr)
      \ee^{-V_z(\eta)}
    \right\rangle_{\gamma_z}
  }{
    D_z
  },
  \qquad
  D_z=\left\langle\ee^{-V_z(\eta)}\right\rangle_{\gamma_z}.
  \label{eq:instanton-physical-source-fluctuation-quotient}
\end{equation}
The exact identity
\begin{equation}
  \mathcal R_{\rm fluc}-1
  =
  \langle U_z\rangle_{\gamma_z}
  +
  \frac{
    \left\langle
    \bigl(U_z-\langle U_z\rangle_{\gamma_z}\bigr)
    \bigl(\ee^{-V_z}-D_z\bigr)
    \right\rangle_{\gamma_z}
  }{D_z}
  \label{eq:instanton-physical-source-fluctuation-covariance}
\end{equation}
separates the Gaussian mean of the selected source from its covariance with
the same nonzero-mode interaction weight that produced the determinant
normalization.  A source-independent determinant calibration fixes \(D_z\).
It does not fix either term in
\eqref{eq:instanton-physical-source-fluctuation-covariance}.

If the retained source variation begins as
\begin{equation}
  U_z(\eta;\mathcal J)
  =
  \ell_a(z;\mathcal J)\eta_a
  +\frac12Q_{ab}(z;\mathcal J)\eta_a\eta_b
  +U_{\ge3,z}(\eta;\mathcal J),
  \qquad
  C_{ab}(z)=\langle\eta_a\eta_b\rangle_{\gamma_z},
  \label{eq:instanton-physical-gaussian-source-expansion}
\end{equation}
then the Gaussian source mean is
\[
  \langle U_z\rangle_{\gamma_z}
  =
  \frac12\operatorname{Tr}(Q_zC_z)
  +\langle U_{\ge3,z}\rangle_{\gamma_z},
\]
because the linear normal-fluctuation term has zero Gaussian mean.  Suppose
on the retained source and size window
\[
  |D_z|\ge d_0,\qquad
  |\langle U_{\ge3,z}\rangle_{\gamma_z}|\le\epsilon_U,\qquad
  \|U_z-\langle U_z\rangle_{\gamma_z}\|_{L^2(\gamma_z)}\le M_U,
  \qquad
  \|\ee^{-V_z}-D_z\|_{L^2(\gamma_z)}\le M_V .
\]
Then
\begin{equation}
  \left|
    \mathcal R_{\rm fluc}(z;\mathcal J)
    -
    1
    -
    \frac12\operatorname{Tr}(Q_zC_z)
  \right|
  \le
  \epsilon_U+\frac{M_UM_V}{d_0}.
  \label{eq:instanton-physical-source-fluctuation-bound}
\end{equation}
After multiplying the hard kernel in
\eqref{eq:instanton-hard-four-source-coefficient}, the contribution of this
normal-fluctuation slot to the absolute window budget is bounded by the
absolute leading kernel mass times the right side of
\eqref{eq:instanton-physical-source-fluctuation-bound}.  If
\(\mathcal Z_z(0;\mathcal J)\) loses rank, the relative quotient
\eqref{eq:instanton-physical-source-fluctuation-quotient} is the wrong
coordinate; the fluctuation contribution must be kept as an absolute source
residual in \eqref{eq:instanton-physical-channel-residual-budget}.
\end{controlledapproximation}

\begin{controlledapproximation}[First source cumulant from normal modes]
\label{ca:instanton-first-source-cumulant-normal-modes}
The term hidden most easily in a determinant-only treatment is the first
source cumulant.  Keep the same zero-mode-deleted Gaussian covariance
\(C_{ab}\), and suppose the normal-mode source variation and interaction begin
as
\[
  U=\ell_a\eta_a+\frac12Q_{ab}\eta_a\eta_b+O(\eta^3),
  \qquad
  V_3=\frac{1}{3!}T_{abc}\eta_a\eta_b\eta_c .
\]
The tensor \(T\) is the cubic variation of the regulated action in the
horizontal normal coordinates.  To first source-dependent order in this
finite chart,
\begin{equation}
\begin{aligned}
  \mathcal R_{\rm fluc}-1
  &=
  \frac12Q_{ab}C_{ab}
  -
  \frac{1}{3!}\ell_iT_{abc}
  \bigl(
    C_{ia}C_{bc}+C_{ib}C_{ac}+C_{ic}C_{ab}
  \bigr)
  +R_{\rm cum}                                      \\
  &=
  \frac12\operatorname{Tr}(QC)
  -
  \frac12\ell_iT_{abc}C_{ia}C_{bc}
  +R_{\rm cum},
\end{aligned}
  \label{eq:instanton-first-source-cumulant-normal}
\end{equation}
where the second line uses the symmetry of the cubic action tensor.  Thus the
linear source deformation \(\ell_a\eta_a\), although it has zero Gaussian
mean, becomes a physical amplitude correction through its covariance with the
cubic fluctuation action.  It is source-dependent amplitude data beyond the
vacuum determinant normalization \(D_z\).

If the retained cumulant remainder obeys
\(|R_{\rm cum}(z;\mathcal J)|\le B_{\rm cum}(z;\mathcal J)\), then its
contribution to the hard source window is controlled by
\[
  \int_{\mathcal W}
  |K_0(z;\mathcal J)|\,B_{\rm cum}(z;\mathcal J)\,\dd z .
\]
The negative controls are again physical: setting \(T=0\) because the
determinant has already been normalized drops a real source response; using
unprojected normal coordinates mixes this cumulant with gauge zero modes; and
absorbing the term into \(C_{\mathcal S}\) turns a source-dependent correction
into a false universal prefactor.
\end{controlledapproximation}

\begin{controlledapproximation}[Normal propagator insertion]
\label{ca:instanton-normal-propagator-source-insertion}
The determinant and the source cumulants still do not exhaust the fluctuation
calculation.  A channel with an additional chirality-preserving spectator
pair, or with a bilinear probe of the gauge or matter fluctuation, asks for a
matrix element of the zero-mode-deleted Green operator in the instanton
background.  Fix a retained instanton chart point \(z=(a,\rho,\Omega)\), and
let \(P_z^\perp\) be the orthogonal projection away from collective, gauge,
and fermionic zero modes in the finite regulator.  If \(K_z\) is the
quadratic normal operator and \(G_z'=P_z^\perp K_z^{-1}P_z^\perp\), define the
source-projected normal propagator coordinate by
\begin{equation}
  \Pi_z^{\rm nz}(J_{\rm out},J_{\rm in})
  =
  \left\langle
    P_z^\perp j_{\rm out}^{\rm amp},
    G_z'\,
    P_z^\perp j_{\rm in}^{\rm amp}
  \right\rangle_z .
  \label{eq:instanton-normal-propagator-projection}
\end{equation}
The amputated source coordinate is
\[
  j^{\rm amp}=S^{-1}L^{-1}j^{\rm phys},
\]
where \(S\) is the finite pole-overlap matrix and \(L\) is the residue
factor in the same source basis used for the zero-mode slots.  The prime in
\(G_z'\) fixes the channel coordinate: an unprimed inverse introduces
arbitrary zero-mode regulator masses and changes the channel.

For a hard zero-mode source block \(\mathcal Z_{\rm zm}^0(z;\mathcal J)\),
the corresponding one-instanton coefficient has the retained form
\begin{equation}
  \mathcal A_{\rm zm+nz}^{I}
  =
  \ee^{\ii\theta}
  \int_{\mathcal W}
  D_z^{I}\,
  \mathcal Z_{\rm zm}^0(z;\mathcal J)\,
  \Pi_z^{\rm nz}(J_{\rm out},J_{\rm in})\,
  \dd z
  +R_{\rm zm+nz}.
  \label{eq:instanton-zero-mode-normal-propagator-channel}
\end{equation}
This is the finite-regulator version of the step in the original instanton
amplitude calculation where fluctuations around the saddle are integrated
along with the collective coordinates.  The density \(D_z^{I}\) and the
zero-mode source determinant are multiplicative weights.  The physical
content of the spectator pair is the bilinear matrix element
\(\Pi_z^{\rm nz}\), which depends on the external sources, amputation
convention, and projection.  A determinant trace such as
\(\operatorname{Tr}'(G_zR_z)\) is a vacuum-response coordinate; it is not the
same datum as \(\langle j_{\rm out},G_z'j_{\rm in}\rangle\).

Assume, on the retained window, that
\[
  D_z=D_z^0(1+\delta_D),\qquad
  \mathcal Z_{\rm zm}=\mathcal Z_{\rm zm}^0(1+\delta_Z),
  \qquad
  |\delta_D|\le e_D,\quad |\delta_Z|\le e_Z,
\]
and that the amputated normal propagator projection satisfies
\[
  \Pi_z^{\rm nz}
  =
  \Pi_z^{0}+\Delta\Pi_z,\qquad
  |\Delta\Pi_z|\le B_\Pi(z).
\]
Then the absolute channel error obeys
\begin{equation}
\begin{aligned}
  |R_{\rm zm+nz}|
  &\le
  \int_{\mathcal W}
  |D_z^0\mathcal Z_{\rm zm}^0|
  \left[
    (1+e_D)(1+e_Z)B_\Pi(z)
    +|\Pi_z^0|
      \bigl((1+e_D)(1+e_Z)-1\bigr)
  \right]\dd z                                      \\
  &\qquad{}+B_{\rm amp}+B_{\rm proj}+B_{\rm sec}.
  \label{eq:instanton-normal-propagator-channel-bound}
\end{aligned}
\end{equation}
The extra terms record pole-overlap inversion, final physical projection, and
sector-isolation errors.  A moduli-space density times a determinant constant
cannot supply \(\Pi_z^{\rm nz}\).  A diagonal residue division cannot replace
the full source-overlap inverse.  A trace over \(G_z'\) loses the external
source vectors.  If the selected zero-mode determinant loses rank, the whole
relative channel coordinate in
\eqref{eq:instanton-zero-mode-normal-propagator-channel} must again be
treated as an absolute residual rather than as a small multiplicative
correction.
\end{controlledapproximation}

\begin{controlledapproximation}[Subtracted normal Green function in a source
channel]
\label{ca:instanton-subtracted-normal-green-matching}
The primed inverse in
\eqref{eq:instanton-normal-propagator-projection} is a distributional Green
kernel with the zero modes removed.  Before subtracting anything one must
state which source/operator class is being computed.  For smooth compact
external sources \(\mathcal U,\mathcal V\), the smeared pairing
\[
  \mathfrak G_{\mathcal R}^{\rm sm}
  (z;\mathcal U,\mathcal V)
  =
  \langle \mathcal U_z,P_z^\perp G_{z,\mathcal R}P_z^\perp
  \mathcal V_z\rangle
\]
is already the source bilinear in that regulator, after the same amputation
and zero-mode projection as in
\eqref{eq:instanton-normal-propagator-projection}.  A free Green function is
not subtracted from this object unless the observable has been declared to be a
background-minus-vacuum difference.  Otherwise the subtraction removes
ordinary physical propagation.

Additional UV matching data enter when the source coordinate is local:
for example a point-supported hard source, a coincident composite operator, or
the local part of a Wilsonian source product.  In that case fix a regulator
\(\mathcal R\), a local diagonal parametrix or declared hard-source
subtraction \(G_{0,z}^{\chi}\), and the heat-kernel coefficient \(A_{1,z}\) in
the same horizontal normal bundle.  The finite local source coordinate is
\begin{equation}
\begin{aligned}
  \mathfrak G_{\mathcal R}^{\rm sub}
  (z;\mathcal U,\mathcal V;\mu)
  &=
  \langle \mathcal U_z,P_z^\perp G_{z,\mathcal R}P_z^\perp\mathcal V_z\rangle
  -\langle \mathcal U_z,G_{0,z}^{\chi}\mathcal V_z\rangle          \\
  &\quad
  -\log(\Lambda/\mu)\,
    \langle \mathcal U_z,A_{1,z}\mathcal V_z\rangle
  -\langle \mathcal U_z,L_{\mathcal R,z}(\mu)\mathcal V_z\rangle .
\end{aligned}
  \label{eq:instanton-subtracted-normal-green-coordinate}
\end{equation}
Here \(L_{\mathcal R,z}(\mu)\) is the finite local counterterm matrix which
matches the local source bilinear to the chosen renormalized operator
coordinate.  The brackets in
\eqref{eq:instanton-subtracted-normal-green-coordinate} are the local pullback
or declared hard-source pairing, not an instruction to subtract full
nonlocal free propagation from an arbitrary smooth source bilinear.
The same projector \(P_z^\perp\) used in the determinant must be used here;
otherwise the subtraction leaves arbitrary zero-mode regulator masses in the
external-source amplitude.

Changing the finite subtraction by
\[
  L_{\mathcal R',z}(\mu)
  =
  L_{\mathcal R,z}(\mu)+\Delta L_z
\]
does not change a completed amplitude.  It changes the Green coordinate by
\(-\langle\mathcal U_z,\Delta L_z\mathcal V_z\rangle\), while the local
Wilsonian coefficient which multiplies the same zero-mode source block
changes by the opposite amount.  Thus the matched channel contribution
\begin{equation}
  \mathcal Z_{\rm zm}^0(z;\mathcal J)
  \bigl[
    \mathfrak G_{\mathcal R}^{\rm sub}
    (z;\mathcal U,\mathcal V;\mu)
    +\mathfrak C_{\mathcal R}^{\rm loc}
    (z;\mathcal U,\mathcal V;\mu)
  \bigr]
  \label{eq:instanton-subtracted-normal-green-matched-channel}
\end{equation}
is invariant under the finite local change.  The determinant identity
\(\delta\log\det'K=\operatorname{Tr}(G'\delta K)\) calibrates the trace response of the
Gaussian normalization; it does not replace
\eqref{eq:instanton-subtracted-normal-green-coordinate}, which is a bilinear
between the actual external source vectors.

If the local projector, parametrix, heat-kernel coefficient, finite local
matching, and source amputation have errors bounded by
\(E_\perp,E_\chi^{\rm loc},E_{1}^{\rm loc},E_L^{\rm loc},E_{\rm amp}\), and
if \(\|P_z^\perp G_{z,\mathcal R}P_z^\perp\|\le M_G\) on the retained normal
space, then on a retained window
\begin{equation}
  |R_{\mathfrak G}|
  \le
  |\mathcal Z_{\rm zm}^0|
  \bigl[
    2M_G E_\perp\|\mathcal U\|\,\|\mathcal V\|
    +E_\chi^{\rm loc}
    +|\log(\Lambda/\mu)|E_1^{\rm loc}
    +E_L^{\rm loc}
  \bigr]
  +E_{\rm amp}.
  \label{eq:instanton-subtracted-normal-green-residual-bound}
\end{equation}
For a smooth-source coordinate the local terms
\(E_\chi^{\rm loc},E_1^{\rm loc},E_L^{\rm loc}\) are absent; for a local
composite coordinate they are load-bearing.  A free subtraction applied to the
wrong source class, a subtraction in a different source basis, a determinant
trace in place of the bilinear, or a finite counterterm shift applied without
the compensating local coefficient gives a different amplitude coordinate.
This is one of the concrete steps in the original instanton calculation where
the fluctuation operator around the saddle, not just its determinant, enters
the physical four-point kernel.
\end{controlledapproximation}

\begin{controlledapproximation}[Spectral cutoff test for local Green matching]
\label{ca:instanton-spectral-local-green-matching-test}
The preceding distinction can be tested without hiding the ultraviolet issue
inside notation.  Fix one zero-mode-deleted normal operator \(K_z\) in
background gauge, acting on the horizontal normal space with positive
eigenpairs
\[
  K_ze_{j,z}=\lambda_{j,z}e_{j,z},
  \qquad
  0<\lambda_{1,z}\le\lambda_{2,z}\le\cdots .
\]
For smooth source vectors with coefficients
\(U_j=\langle e_{j,z},\mathcal U_z\rangle\) and
\(V_j=\langle e_{j,z},\mathcal V_z\rangle\), the cutoff pairing is
\begin{equation}
  \mathfrak G_{N}^{\rm sm}(z;\mathcal U,\mathcal V)
  =
  \sum_{j\le N}\frac{\overline{U_j}V_j}{\lambda_{j,z}} .
  \label{eq:instanton-spectral-smooth-green-pairing}
\end{equation}
If the tail
\(\sum_{j>N}|\overline{U_j}V_j|/\lambda_{j,z}\) is small in the chosen source
norm, this is the regulated source bilinear.  There is no \(A_0N\) or
\(A_1H_N\) term to subtract from it; imposing such a subtraction changes the
observable.

A local source coordinate is different.  Let \(\ell_{x,z}\) be a declared
local pullback functional, such as the diagonal value of a hard-source or
composite-operator coordinate after the spin, color, and density conventions
have been fixed.  Group the nonzero modes into regulator shells \(S_n\), and
suppose the local diagonal shell has the expansion
\begin{equation}
  \sum_{j\in S_n}
  \frac{\overline{\ell_{x,z}(e_{j,z})}\ell_{x,z}(e_{j,z})}
       {\lambda_{j,z}}
  =
  A_{0,z}(x)+\frac{A_{1,z}(x)}{n}+r_{n,z}(x),
  \qquad
  \sum_{n\ge1}|r_{n,z}(x)|<\infty .
  \label{eq:instanton-spectral-local-shell-expansion}
\end{equation}
The cutoff local coordinate is then
\[
  \mathfrak G_N^{\rm loc}(z;x)
  =
  A_{0,z}(x)N+A_{1,z}(x)H_N
  +\sum_{n\le N}r_{n,z}(x),
  \qquad
  H_N=\sum_{n=1}^{N}\frac1n .
\]
In this discrete regulator the matched local coordinate is
\begin{equation}
  \mathfrak G_{N,\mu}^{\rm ren}(z;x)
  =
  \mathfrak G_N^{\rm loc}(z;x)
  -A_{0,z}(x)N-A_{1,z}(x)H_N
  -L_{N,z}(x;\mu).
  \label{eq:instanton-spectral-renormalized-local-green}
\end{equation}
Passing from \(H_N\) to \(\log(\Lambda/\mu)\) is a finite scheme change
absorbed into \(L_{N,z}\).  The coefficients \(A_0,A_1\) are local
heat-kernel or operator-mixing data of the chosen \(\ell_{x,z}\), not
parameters inferred from the final finite answer.  With coefficient errors
\(\delta A_0,\delta A_1,\delta L\), the retained residual obeys
\begin{equation}
  |R_N^{\rm loc}|
  \le
  N|\delta A_0|+H_N|\delta A_1|+|\delta L|
  +\sum_{n>N}|r_{n,z}(x)|.
  \label{eq:instanton-spectral-local-green-bound}
\end{equation}
Thus the same spectral regulator gives two different lessons.  Smooth
external sources test propagation through the primed Green operator.  Local or
coincident source coordinates test the heat-kernel subtraction and Wilsonian
operator basis.  A four-point instanton amplitude has to say which of these it
uses before any fluctuation determinant, source quotient, or physical
projection can be interpreted.
\end{controlledapproximation}

Thus a hard instanton amplitude has several separate fluctuation entries.  The
universal logarithmic determinant supplies the \((\mu\rho)^{b_0}\) density
factor.  The finite determinant constant fixes a scheme-dependent Gaussian
normalization.  The primed resolvent response
\eqref{eq:instanton-primed-boson-determinant-response}--\eqref{eq:instanton-primed-fermion-determinant-response}
records how the determinant itself changes under the selected source or
background perturbation.  The quotient
\eqref{eq:instanton-physical-source-fluctuation-quotient} tests the selected
source under that Gaussian and its interaction remainders, including cumulants
such as \eqref{eq:instanton-first-source-cumulant-normal}.  The
source-projected normal propagator
\eqref{eq:instanton-normal-propagator-projection}, after the subtraction and
local matching in
\eqref{eq:instanton-subtracted-normal-green-coordinate} and the
source-class test in
\eqref{eq:instanton-spectral-renormalized-local-green}, supplies a further
external-state-dependent fluctuation coordinate when the channel probes
nonzero-mode propagation.  Collapsing these entries into a single ``measure
prefactor'' loses the amplitude calculation.

\section{The Hard Channel As An Assembled Amplitude}
\label{sec:instanton-hard-amplitude-assembly}

The preceding ingredients are useful only when they are recombined in the same
observable coordinate.  For the \(SU(3)\), \(N_f=2\), hard four-source channel,
let \(s=Q\rho\), let \(\Omega\) denote the retained color-orientation
coordinate, and let \(\chi_{\rm phys}\) record the declared source window,
amputation convention, and physical projection after the center
momentum-conservation projection has been applied.  A finite-regulator assembly has
the form
\begin{equation}
\begin{aligned}
  \mathcal A_{4,\Lambda}^{\rm hard}(Q;\mathcal J)
  &=
  \ee^{\ii\theta}
  C_{\mathcal S}\,
  \Gamma_{\rm coll}(Q)\,
  \Lambda_{\rm ht}^{29/3}Q^{-35/3}
  \int_{\mathcal W}
  K_0(s,\Omega;\mathcal J)\,
  \mathcal R_{\rm det}(s,\Omega)\,                       \\
  &\qquad\qquad{}\times
  \mathcal R_{\rm fluc}(s,\Omega;\mathcal J)\,
  \mathcal R_{\rm zm}(s,\Omega;\mathcal J)\,
  \mathcal R_{\rm phys}(s,\Omega;\mathcal J)\,
  \dd s\,\dd\Omega
  +R_{\rm ext}(Q;\mathcal J),
\end{aligned}
  \label{eq:instanton-hard-assembled-channel}
\end{equation}
where
\[
\begin{aligned}
  K_0(s,\Omega;\mathcal J)
  &=
  s^{32/3}
  \mathcal P_{\rm Haar}(\Omega)\,
  \det B_R^0(s,\Omega;\mathcal J)\,
  \det B_L^0(s,\Omega;\mathcal J)              \\
  &\qquad{}\times
  \prod_{\ell=1}^4F_\ell(c_\ell s)\,
  \chi_{\rm phys}^{0}(s,\Omega;\mathcal J).
\end{aligned}
\]
The constant \(C_{\mathcal S}\) is the finite determinant and orientation
normalization in the chosen scheme.  The logarithmic determinant and running
action have been traded for \(\Lambda_{\rm ht}^{29/3}Q^{-35/3}\), while the
bosonic collective-coordinate normalization remains as the running
\(\Gamma_{\rm coll}(Q)\) factor.  The remaining factors have distinct jobs:
\(\mathcal R_{\rm det}\) is the finite determinant-shape mismatch after
scheme matching, \(\mathcal R_{\rm fluc}\) is the source-dependent quotient in
\eqref{eq:instanton-physical-source-fluctuation-quotient},
\(\mathcal R_{\rm zm}\) is the zero-mode/source-frame stability factor, and
\(\mathcal R_{\rm phys}\) is the finite projection or continuation mismatch.
The term \(R_{\rm ext}\) contains sector isolation, infrared, cut, endpoint,
and continuum errors not represented by the retained window.

\begin{controlledapproximation}[Hard amplitude assembly control]
\label{ca:instanton-hard-amplitude-assembly-ledger}
Let
\begin{equation}
  H_{\mathcal C}^{0}(\mathcal J)
  =
  \int_{\mathcal W}K_0(s,\Omega;\mathcal J)\,\dd s\,\dd\Omega,
  \qquad
  M_{\mathcal C}^{0}(\mathcal J)
  =
  \int_{\mathcal W}|K_0(s,\Omega;\mathcal J)|\,\dd s\,\dd\Omega .
  \label{eq:instanton-hard-assembly-leading-window}
\end{equation}
Assume on the retained window that
\[
  \mathcal R_x=1+\delta_x,\qquad
  |\delta_x|\le e_x,
  \qquad x\in\{\det,{\rm fluc},{\rm zm},{\rm phys}\}.
\]
Then the leading assembled coefficient
\[
  \mathcal A_{4}^{0}(Q;\mathcal J)
  =
  \ee^{\ii\theta}
  C_{\mathcal S}
  \Gamma_{\rm coll}(Q)
  \Lambda_{\rm ht}^{29/3}Q^{-35/3}
  H_{\mathcal C}^{0}(\mathcal J)
\]
has the absolute bound
\begin{equation}
\begin{aligned}
  \left|
    \mathcal A_{4,\Lambda}^{\rm hard}(Q;\mathcal J)
    -
    \mathcal A_4^0(Q;\mathcal J)
  \right|
  &\le
  |C_{\mathcal S}|\Gamma_{\rm coll}(Q)
  \Lambda_{\rm ht}^{29/3}Q^{-35/3}
  \epsilon_{\rm asm}\,
  M_{\mathcal C}^{0}(\mathcal J)
  +|R_{\rm ext}(Q;\mathcal J)|,                         \\
  \epsilon_{\rm asm}
  &=
  (1+e_{\det})(1+e_{\rm fluc})(1+e_{\rm zm})(1+e_{\rm phys})-1 .
\end{aligned}
  \label{eq:instanton-hard-assembly-error-bound}
\end{equation}
If a relative statement is wanted, a noncancellation margin must be supplied:
\[
  |H_{\mathcal C}^{0}(\mathcal J)|
  \ge h_0 M_{\mathcal C}^{0}(\mathcal J),
  \qquad h_0>0 .
\]
Only then does \eqref{eq:instanton-hard-assembly-error-bound} imply a
relative correction of size at most
\[
  \frac{\epsilon_{\rm asm}}{h_0}
  +
  \frac{|R_{\rm ext}(Q;\mathcal J)|}{
    |C_{\mathcal S}|\Gamma_{\rm coll}(Q)
    \Lambda_{\rm ht}^{29/3}Q^{-35/3}
    |H_{\mathcal C}^{0}(\mathcal J)|
  } .
\]
Without this margin the correct control is absolute, not relative; a signed
window cancellation can make a tiny leading number even when the source,
determinant, and projection errors are individually small.
\end{controlledapproximation}

The finite determinant constant in
\eqref{eq:instanton-hard-assembled-channel} may be computed directly in a
declared regulator, as in the density calculation of
Chapter~\ref{ch:volume-ii-20}.  It may also be fixed by one reference
amplitude.  That calibration is useful only if it is kept in its proper
coordinate: it fixes one source-independent constant, not the source
fluctuation quotient, not zero-mode rank, and not the final physical
projection.

\begin{controlledapproximation}[Reference-channel determinant calibration]
\label{ca:instanton-hard-reference-channel-calibration}
Fix the determinant scheme, source frame, amputation convention, hard-scale
variable, and physical projection class.  Let
\[
  \mathfrak G(Q)
  =
  \ee^{\ii\theta}
  \Gamma_{\rm coll}(Q)\Lambda_{\rm ht}^{29/3}Q^{-35/3}.
\]
For a finite family of source channels \(\mathcal C_\alpha\), suppose the
retained leading integrals have been reduced to
\begin{equation}
  B_\alpha
  =
  \int_{\mathcal W}
  K_\alpha^{0}(s,\Omega)\,\dd s\,\dd\Omega,
  \qquad
  M_\alpha
  =
  \int_{\mathcal W}
  |K_\alpha^{0}(s,\Omega)|\,\dd s\,\dd\Omega,
  \label{eq:instanton-hard-reference-channel-integrals}
\end{equation}
where \(K_\alpha^{0}\) includes the center constraint, Haar projection,
zero-mode overlap determinants, declared leading source window, amputation
residues, and leading physical projection for that channel.  All
source-dependent fluctuation quotients, determinant-shape mismatches,
endpoint tails not retained in \(K_\alpha^{0}\), and bridge errors are put in
\(\Delta_\alpha\).  If
\begin{equation}
  A_\alpha^{\rm phys}
  =
  \mathfrak G(Q) C_{\mathcal S}B_\alpha+\Delta_\alpha,
  \qquad
  |\Delta_\alpha|\le E_\alpha,
  \label{eq:instanton-hard-reference-family}
\end{equation}
and a reference channel satisfies
\begin{equation}
  A_0^{\rm phys}
  =
  \mathfrak G(Q) C_{\mathcal S}B_0+\Delta_0,
  \qquad
  B_0\ne0,\qquad |\Delta_0|\le E_0,
  \label{eq:instanton-hard-reference-input}
\end{equation}
then the calibrated prediction for a target channel is
\begin{equation}
  \widehat A_\alpha^{\rm phys}
  =
  A_0^{\rm phys}\frac{B_\alpha}{B_0},
  \label{eq:instanton-hard-reference-prediction}
\end{equation}
and the exact residual identity is
\begin{equation}
  A_\alpha^{\rm phys}-\widehat A_\alpha^{\rm phys}
  =
  \Delta_\alpha-\frac{B_\alpha}{B_0}\Delta_0,
  \qquad
  \left|
  A_\alpha^{\rm phys}-\widehat A_\alpha^{\rm phys}
  \right|
  \le
  E_\alpha+\left|\frac{B_\alpha}{B_0}\right|E_0 .
  \label{eq:instanton-hard-reference-error}
\end{equation}
The ratio is usable only with a reference noncancellation margin
\begin{equation}
  \kappa_0=\frac{|B_0|}{M_0}>0,
  \qquad
  \left|\frac{B_\alpha}{B_0}\right|
  \le
  \frac{M_\alpha}{\kappa_0M_0}.
  \label{eq:instanton-hard-reference-margin}
\end{equation}

The calibration has limited scope.  A reference channel with rank loss, a
vanishing Haar projection, or a signed cancellation which makes \(\kappa_0\)
tiny is a poor normalization coordinate.  If the target channel has a
different source-fluctuation quotient or a different pole, spectral, or OPE
projection and those factors have not been included in \(B_\alpha\), they
remain in \(\Delta_\alpha\) outside the calibrated constant \(C_{\mathcal S}\).
A reference amplitude therefore transports one finite determinant
normalization across a declared same-frame source family; the Euclidean source
kernel still reaches a physical observable only through the projection bridge
below.
\end{controlledapproximation}

\begin{remark}[Finite determinant scheme-transport architecture]
\label{rem:instanton-finite-determinant-scheme-transport-architecture}
The universal finite one-loop determinant and orientation constant is not an
invariant number by itself, but its transformation is narrower than a full
amplitude transformation.  It belongs to the regulator, coupling coordinate,
bosonic zero-mode normalization, and orientation-measure convention.  It does
not absorb the source basis or the physical projection.  Compare two
determinant descriptions \({\mathcal S}\) and \({\mathcal S}'\) at the same hard
scale \(Q\), and write
\[
  X_{\mathcal S}(Q)=\frac{8\pi^2}{g_{\mathcal S}^2(Q)},
  \qquad
  \Gamma_{\mathcal S}(Q)=X_{\mathcal S}(Q)^{2N_c}.
\]
Suppose the universal density factors are related by
\[
  \ee^{-X_{{\mathcal S}'}}=\lambda_{\rm act}^{-1}\ee^{-X_{\mathcal S}},
  \qquad
  \dd\Omega_{{\mathcal S}'}=\lambda_{\Omega}\dd\Omega_{\mathcal S},
\]
\[
  C_{{\mathcal S}'}^{\rm inst}
  \Gamma_{{\mathcal S}'}(Q)
  \lambda_{\rm act}^{-1}
  \lambda_{\Omega}
  =
  C_{\mathcal S}^{\rm inst}\Gamma_{\mathcal S}(Q).
\]
Equivalently,
\begin{equation}
  C_{{\mathcal S}'}^{\rm inst}
  =
  C_{\mathcal S}^{\rm inst}\,
  \lambda_{\rm act}
  \frac{\Gamma_{\mathcal S}(Q)}{\Gamma_{{\mathcal S}'}(Q)}
  \lambda_{\Omega}^{-1}.
  \label{eq:instanton-finite-determinant-constant-transform}
\end{equation}
No source determinant or projection factor appears in
\eqref{eq:instanton-finite-determinant-constant-transform}.  Those factors act
on the channel vector.  For a finite family of retained operators, let
\(\mathbf c_{\mathcal S}\) be the row of zero-mode/source coefficients and
\(\mathbf m_{\mathcal S}\) the column of physical matrix elements or projection
data.  If the operator/source basis is changed by
\(\mathbf O_{{\mathcal S}'}=T\mathbf O_{\mathcal S}\), then
\begin{equation}
  \mathbf c_{{\mathcal S}'}=\mathbf c_{\mathcal S}T^{-1},
  \qquad
  \mathbf m_{{\mathcal S}'}=T\mathbf m_{\mathcal S},
  \qquad
  \mathbf c_{{\mathcal S}'}\mathbf m_{{\mathcal S}'}
  =
  \mathbf c_{\mathcal S}\mathbf m_{\mathcal S}.
  \label{eq:instanton-source-projection-covariance}
\end{equation}
The one-channel shorthand
\[
  \det B_R^{{\mathcal S}'}\det B_L^{{\mathcal S}'}
  =
  \lambda_{\rm src}\det B_R^{\mathcal S}\det B_L^{\mathcal S},
  \qquad
  \mathfrak L_{\rm phys}^{{\mathcal S}'}
  =
  \lambda_{\rm phys}\mathfrak L_{\rm phys}^{\mathcal S}.
\]
is therefore a statement about a full channel coefficient, not about the
universal determinant constant.  In more than one channel the two displayed
scalars are matrices or covectors, and absorbing them into
\(C_{\mathcal S}^{\rm inst}\) would make a source-independent determinant
number depend on the channel chosen for calibration.

If only the coupling coordinate changes and
\(X_{{\mathcal S}'}=X_{\mathcal S}+\kappa+O(g^2)\), then
\(\lambda_{\rm act}=\ee^\kappa+O(g^2)\) and
\(C_{{\mathcal S}'}^{\rm inst}=\ee^\kappa C_{\mathcal S}^{\rm inst}(1+O(g^2))\)
at one-loop accuracy when the orientation convention is unchanged.  The exact
finite-coordinate comparison, however, keeps the
\(\Gamma_{\mathcal S}/\Gamma_{{\mathcal S}'}\) factor in
\eqref{eq:instanton-finite-determinant-constant-transform}.  This is why a
quoted Pauli--Villars or subtraction-scheme constant cannot be moved into a
hard amplitude unless the coupling coordinate, bosonic zero-mode power, and
orientation normalization have been moved, and unless the separate
source/projection covariance \eqref{eq:instanton-source-projection-covariance}
has been checked for the channel being projected.

This paragraph is an architecture identity, not a determinant conversion by
itself.  The numbers \(\lambda_{\rm act}\), \(\lambda_\Omega\), and
\(\Gamma_{\mathcal S}/\Gamma_{{\mathcal S}'}\) have evidentiary force only
after the two regulated fluctuation measures have both been computed.  A
companion calculation which chooses these multipliers and then defines
\(C_{{\mathcal S}'}^{\rm inst}\) from
\eqref{eq:instanton-finite-determinant-constant-transform} tests covariance
typing, but it has not computed a Pauli--Villars-to-\(\overline{\rm MS}\), or
any other named continuum, determinant conversion.

With imperfect universal transport, write every determinant-density conversion
factor as its declared value times \(1+\delta_y\), with
\(|\delta_y|\le e_y<1\) for
\[
  y\in\{\Gamma,{\rm act},\Omega\}.
\]
Let \(B_{\mathcal C}^{\mathcal S}=\mathbf c_{\mathcal S}\mathbf m_{\mathcal S}\)
denote the retained channel kernel, including the declared source window and
physical projection, and suppose the source/projection/shape transport obeys
\[
  \left|
  B_{\mathcal C}^{{\mathcal S}'}-B_{\mathcal C}^{\mathcal S}
  \right|
  \le E_{\mathcal C}^{\rm ch}.
\]
Then the same leading-window amplitude has the residual bound
\begin{equation}
  \begin{aligned}
  \epsilon_{\rm sch}
  =
  \prod_y(1+e_y)-1,
  \qquad
  |\Delta A_{\rm sch}|
  \le
  |C_{\mathcal S}^{\rm inst}\Gamma_{\mathcal S}
  B_{\mathcal C}^{\mathcal S}|\,
  \epsilon_{\rm sch}
  +
  |C_{\mathcal S}^{\rm inst}\Gamma_{\mathcal S}|\,
  E_{\mathcal C}^{\rm ch}
  +R_{\rm bridge}.
  \end{aligned}
  \label{eq:instanton-finite-determinant-scheme-residual}
\end{equation}
The bridge term contains source-fluctuation, endpoint, sector, and continuation
mismatches not represented by the finite channel vector.  Thus scheme transport
is still an amplitude calculation: moving only the determinant constant, only
the running coupling, or a source frame without its contragredient projection
changes the QFT coordinate being compared.
\end{remark}

\begin{controlledapproximation}[Finite determinant conversion benchmark]
\label{ca:instanton-finite-determinant-conversion-benchmark}
The finite-regulator test required by the preceding architecture starts before
any source projection.  For a determinant regulator \(\mathcal R\), define the
retained universal density by
\begin{equation}
  \mathcal D_{\mathcal R}(Q)
  =
  a_{\mathcal R}\,
  x_{\mathcal R}^{2N_c}\,
  \Omega_{\mathcal R}\,
  J_{\mathcal R}^{0}\,
  \frac{
    \det{}'D_{F,\mathcal R}\,
    \det{}'M_{{\rm gh},\mathcal R}
  }{
    \bigl(\det{}'H_{B,\mathcal R}\bigr)^{1/2}
  },
  \qquad
  a_{\mathcal R}=\ee^{-X_{\mathcal R}(Q)} .
  \label{eq:instanton-finite-determinant-benchmark-density}
\end{equation}
The factors are read before any channel source is projected: \(a_{\mathcal R}\)
is the finite action weight, \(x_{\mathcal R}^{2N_c}\) is the running
collective-coordinate power, \(\Omega_{\mathcal R}\) is the orientation-volume
normalization, \(J_{\mathcal R}^{0}\) is the bosonic zero-mode Jacobian, and
the determinant ratio is zero-mode deleted in the same gauge slice.

Take \(N_c=2\) and the following two finite regulators:
\[
\begin{array}{c|ccccc}
\mathcal R
& a_{\mathcal R}
& x_{\mathcal R}
& \Omega_{\mathcal R}
& J_{\mathcal R}^{0}
& (\det{}'D_F,\det{}'M_{\rm gh},(\det{}'H_B)^{1/2})
\\ \hline
\mathcal S
& \frac23
& \frac32
& \frac57
& \frac54
& (2,2,4)
\\
\mathcal S'
& \frac35
& \frac43
& \frac78
& 1
& (3,3,5).
\end{array}
\]
Thus
\[
  C_{\mathcal S}^{\rm det}=\frac54,
  \qquad
  C_{{\mathcal S}'}^{\rm det}=\frac95,
  \qquad
  \Gamma_{\mathcal S}=\left(\frac32\right)^4,
  \qquad
  \Gamma_{{\mathcal S}'}=\left(\frac43\right)^4 .
\]
The independently computed density conversion is
\begin{equation}
  K_{\mathcal S\to{\mathcal S}'}
  =
  \frac{\mathcal D_{{\mathcal S}'}}{\mathcal D_{\mathcal S}}
  =
  \frac{50176}{50625}.
  \label{eq:instanton-finite-determinant-benchmark-ratio}
\end{equation}
The architecture factors in
\eqref{eq:instanton-finite-determinant-constant-transform} are
\[
  \lambda_{\rm act}=\frac{10}{9},
  \qquad
  \lambda_\Omega=\frac{49}{40},
  \qquad
  \frac{\Gamma_{\mathcal S}}{\Gamma_{{\mathcal S}'}}
  =
  \frac{6561}{4096}.
\]
They would transport \(C_{\mathcal S}^{\rm det}\) to
\[
  C_{{\mathcal S}'}^{\rm tr}
  =
  \frac{91125}{50176}.
\]
The independently computed determinant constant in \(\mathcal S'\) is
\(9/5\), and therefore
\[
  \frac{C_{{\mathcal S}'}^{\rm det}}{C_{{\mathcal S}'}^{\rm tr}}
  =
  \frac{50176}{50625}
  =
  K_{\mathcal S\to{\mathcal S}'} .
\]
This is a finite conversion calculation, not a tautological definition of
\(C_{{\mathcal S}'}\).  If \(\mathcal S\) and \(\mathcal S'\) are asserted to
represent the same renormalized instanton amplitude in the same source channel,
then the remaining matching factor must be the inverse of
\eqref{eq:instanton-finite-determinant-benchmark-ratio}, or the difference must
be carried as a bridge residual:
\begin{equation}
  \Delta_{\rm bench}
  =
  B_{\mathcal C}
  \bigl(\mathcal D_{{\mathcal S}'}-\mathcal D_{\mathcal S}\bigr).
  \label{eq:instanton-finite-determinant-benchmark-residual}
\end{equation}
Omitting the action weight, running zero-mode power, zero-mode Jacobian,
orientation volume, or zero-mode-deleted determinant ratio changes
\eqref{eq:instanton-finite-determinant-benchmark-ratio}.  A quoted continuum
scheme constant closes the same question only after its own two regulated
measures have been defined at this level of detail.
\end{controlledapproximation}

This assembly formula is the point at which the calculation becomes a
physical amplitude.  Quoting the BPST density tests only the logarithmic
fluctuation normalization.  Quoting the zero-mode determinant tests only chirality and
source rank.  Quoting
\(\mathcal R_{\rm fluc}\) tests only the selected source under the normal
Gaussian.  The hard coefficient is obtained only after these pieces multiply
the same Haar-projected, amputated, windowed, and physically projected kernel.
Changing any of those data changes the amplitude coordinate, even though the
underlying instanton saddle is unchanged.

\section{A Source Kernel Still Needs a Physical Projection}
\label{sec:instanton-source-kernel-physical-projection}

The word ``amplitude'' should not hide the last map.  A finite Euclidean
instanton source kernel
\(\mathcal K_{E,\Lambda}[\mathcal J]\) is a source-functional coefficient in a
regulated path integral.  It becomes a scattering amplitude, an OPE
coefficient, a hadronic matrix element, or a spectral-bin contribution only
after the corresponding projection has been declared.  In a confining gauge
theory this point is especially important: a colored partonic instanton kernel
can be a useful auxiliary object, but it is not by itself an LSZ amplitude
unless a gauge-invariant external source, an infrared deformation, or a
separate partonic matching convention has been supplied.

For a stable gauge-invariant external channel the projection is a pole-window
operation.  If \(O_i,O_f\) are interpolating sources with nonzero overlap with
states \(|i\rangle,|f\rangle\), the large Euclidean-time window has the form
\begin{equation}
  G_{fi}^{I}(T,t)
  =
  Z_f Z_i\,
  \ee^{-E_f(T-t)}\ee^{-E_i t}\,
  \mathcal M_{fi}^{I}
  +R_{\rm exc}^{I}(T,t),
  \label{eq:instanton-hadronic-pole-window}
\end{equation}
after the instanton kernel has been assembled in the source channel.  The
matrix element \(\mathcal M_{fi}^{I}\) is obtained by isolating this pole and
then dividing by \(Z_fZ_i\).  Dividing a raw source correlator by guessed
overlaps before the pole has been isolated leaves excited-state leakage in the
answer, and a vanishing source overlap makes the extraction undefined.

For an inclusive or Euclidean spectral observable the projection is different.
A Euclidean source value has a Stieltjes form
\begin{equation}
  G_E^I(Q^2)
  =
  P_{\rm ct}(Q^2)
  +
  \int_0^\infty \frac{\dd\rho_I(\sigma)}{\sigma+Q^2},
  \qquad
  \mathfrak B_{\Delta}^I
  =
  \int_{\Delta}\dd\rho_I(\sigma),
  \label{eq:instanton-spectral-projection-window}
\end{equation}
where \(P_{\rm ct}\) is a local contact polynomial.  The bin
\(\mathfrak B_{\Delta}^I\) is a discontinuity or spectral functional; it is
not a single spacelike Euclidean value.  Contact polynomials can change
\(G_E^I(Q^2)\) while carrying no separated spectral weight in \(\Delta\), and
two positive spectral measures can agree at one Euclidean point while giving
different bin weights.  A physical spectral claim therefore needs an inversion
or window-stability input, not just a source-kernel number.

\begin{controlledapproximation}[Physical projection bridge for instanton
source kernels]
\label{ca:instanton-source-kernel-physical-projection}
Let \(\Pi_{\rm phys}\) be one declared physical projection: a stable-particle
pole residue, a spectral-bin functional, a discontinuity functional, an OPE
matching map, or an infrared-safe inclusive measurement.  Applying it to the
assembled source
kernel gives a leading physical coordinate
\[
  \mathcal A_{\rm phys}^{0}
  =
  \Pi_{\rm phys}^{0}[\mathcal K_{E,\Lambda}^{0}].
\]
The comparison with the actual physical quantity must budget the bridge
residuals
\[
  \Delta_{\rm phys}
  =
  \Delta_{\rm reg}
  +\Delta_{\rm cont}
  +\Delta_{\rm pole/bin}
  +\Delta_{\rm IR}
  +\Delta_{\rm cut}
  +\Delta_{\rm match}
  +\Delta_{\rm end},
\]
where the entries denote, respectively, regulator transport, analytic
continuation, pole or spectral-bin isolation, infrared completion, unitarity
cut or discontinuity normalization, operator/source matching, and retained
size-window endpoint error.  If
\[
  |\Delta_x|\le B_x
  \qquad
  x\in\{{\rm reg,cont,pole/bin,IR,cut,match,end}\},
\]
then
\begin{equation}
  \left|
    \mathcal A_{\rm phys}
    -
    \Pi_{\rm phys}^{0}[\mathcal K_{E,\Lambda}^{0}]
  \right|
  \le
  B_{\rm reg}+B_{\rm cont}+B_{\rm pole/bin}+B_{\rm IR}
  +B_{\rm cut}+B_{\rm match}+B_{\rm end}.
  \label{eq:instanton-physical-bridge-residual-bound}
\end{equation}

The negative controls are immediate.  A color-nonsinglet auxiliary kernel has
no standalone physical LSZ map in a confining theory.  A pole residue cannot
be read before the source overlap and excited-state window are controlled.  A
spectral bin cannot be inferred from one Euclidean Stieltjes value.  An OPE
coefficient changes when the operator basis or matching scheme is changed.  A
real-time rate requires a retarded/KMS continuation rather than a Euclidean
static source value.  Thus the projection bridge is a separate physical
calculation, not a relabelling of the instanton density or the hard source
kernel.
\end{controlledapproximation}

\begin{controlledapproximation}[Pole-normalized four-source extraction with
source mixing]
\label{ca:instanton-pole-normalized-four-source-extraction}
The pole map in
\eqref{eq:instanton-hadronic-pole-window} is rarely a scalar division by one
source overlap.  Choose finite source bases
\(\{O_i^a\}_{a=1}^{r_i}\) and \(\{O_f^b\}_{b=1}^{r_f}\) for the selected
incoming and outgoing gauge-invariant multiplets.  After the ordinary
two-point functions have fixed the overlap matrices \(Z_i,Z_f\), strip the
known exponential factors from the instanton four-source window and write
\begin{equation}
  \widetilde C_{ba}^{I}(T,t)
  =
  \bigl(Z_f\,M_{fi}^{I}\,Z_i^\dagger\bigr)_{ba}
  +\mathcal E_{ba}^{I}(T,t).
  \label{eq:instanton-pole-matrix-window}
\end{equation}
The matrix \(M_{fi}^{I}\) is the physical instanton contribution to the
selected matrix element in this pole subspace.  It is not the raw Euclidean
kernel \(\widetilde C^{I}\).  When \(Z_f\) and \(Z_i\) are invertible on the
chosen source subspaces, the amputated extraction is
\begin{equation}
  \widehat M_{fi}^{I}
  =
  Z_f^{-1}\,\widetilde C^{I}\,(Z_i^\dagger)^{-1}.
  \label{eq:instanton-pole-amputated-matrix}
\end{equation}
If either overlap matrix loses rank, the chosen source basis does not define
this physical coordinate; one must change sources or keep the statement as a
source-correlator estimate.

Let \(\|\cdot\|_{\max}\) be the entrywise maximum norm and suppose
\[
  \|\mathcal E^{I}\|_{\max}\le\epsilon_{\rm pole},
  \qquad
  \|Z_f^{-1}\|_{\max}\le\kappa_f,
  \qquad
  \|(Z_i^\dagger)^{-1}\|_{\max}\le\kappa_i .
\]
Then
\begin{equation}
  \|
  \widehat M_{fi}^{I}-M_{fi}^{I}
  \|_{\max}
  \le
  r_f r_i\,\kappa_f\kappa_i\,\epsilon_{\rm pole}.
  \label{eq:instanton-pole-amputation-error-bound}
\end{equation}
The factors \(r_f,r_i\) are just the finite matrix-multiplication loss in the
entrywise norm.  They are the finite-dimensional trace of the same physical
fact: source mixing and pole residues amplify excited-state and continuum
leakage.

This block is the pole analogue of the hard-source amputation in
\eqref{eq:instanton-hard-assembled-channel}.  Dividing by only the diagonal
entries of \(Z_i\) and \(Z_f\) gives the wrong result when the source basis
mixes.  Treating \(\widetilde C^{I}\) itself as \(M_{fi}^{I}\) keeps source
normalizations in the answer.  Absorbing \(\mathcal E^{I}\) into the
determinant constant confuses excited-state, continuum, sector, and
projection leakage with the one-loop fluctuation density.  Thus the
four-source instanton coefficient becomes a hadronic matrix element only
after matrix pole isolation, amputation, and the corresponding residual
bound have been supplied.
\end{controlledapproximation}

\begin{controlledapproximation}[Inclusive cut from an instanton amplitude]
\label{ca:instanton-inclusive-cut-quadratic-projection}
The pole-extracted \(Q=1\) coefficient is an amplitude coordinate.  A positive
inclusive cut or rate is a later quadratic projection, with the conjugate
sector and the measurement included.  Let a retained one-instanton channel,
after source differentiation, pole or OPE projection, and amputation, give a
finite vector of physical amplitudes
\[
  a_X^{I}
  =
  \ee^{\ii\theta}
  \mathcal N_I\,\mathfrak a_X ,
  \qquad X\in\Delta ,
\]
where \(\Delta\) labels the selected final-state bin or finite spectral
resolution.  Let \(W_{XY}\) be the positive measurement or bin metric in the
same physical basis.  The corresponding contribution to an inclusive
probability, or to the cut of a neutral forward correlator in this retained
bin, is built by pairing the one-instanton amplitude with its conjugate:
\begin{equation}
  \mathcal Q_{\Delta}^{I\bar I}[W]
  =
  \sum_{X,Y\in\Delta}
  \overline{a_X^{I}}\,
  W_{XY}\,
  a_Y^{I}
  +R_{\rm cut}
  =
  |\mathcal N_I|^2\,
  \mathfrak a^\dagger W\mathfrak a
  +R_{\rm cut}.
  \label{eq:instanton-inclusive-quadratic-cut}
\end{equation}
The \(\theta\)-phase cancels in this positive \(I\bar I\) cut.  The same
one-instanton saddle can still contribute linearly to a chirality-violating
source correlator or to an interference term
\[
  2\Re\,
  \bigl(
    b^\dagger W a^I
  \bigr)
\]
only when the reference amplitude \(b\) lives in the same external-state,
source, and selection-rule coordinate.  Otherwise the relevant interference
matrix element is zero or undefined, not a rate obtained by summing
\(\ee^{\ii\theta}\mathfrak a_X\).

This distinction is part of unitarity normalization.  The optical-theorem
side is a discontinuity or cut of a forward physical object, so it uses the
same external residues, the same final-state measure, and the same
infrared/completeness convention on both sides.  A Euclidean four-source
coefficient, a color-auxiliary kernel, or a pole-amputated vector in the
wrong source basis is not yet the vector \(\mathfrak a\) appearing in
\eqref{eq:instanton-inclusive-quadratic-cut}.

The finite error propagation is also quadratic.  Suppose the retained
amplitude vector is \(\mathfrak a^0+\delta\mathfrak a\), with
\[
  \|\mathfrak a^0\|_{\max}\le M,\qquad
  \|\delta\mathfrak a\|_{\max}\le\epsilon,\qquad
  \max_X\sum_Y |W_{XY}|\le L,\qquad |\Delta|=n .
\]
Then
\begin{equation}
\begin{aligned}
  &\left|
  (\mathfrak a^0+\delta\mathfrak a)^\dagger
  W
  (\mathfrak a^0+\delta\mathfrak a)
  -
  (\mathfrak a^0)^\dagger W\mathfrak a^0
  \right|                                      \\
  &\qquad\le
  2nML\epsilon+nL\epsilon^2 .
\end{aligned}
  \label{eq:instanton-inclusive-quadratic-error-bound}
\end{equation}
Thus the source-amplitude residuals from determinant shape, zero-mode rank,
pole leakage, endpoint tails, and physical projection must be transported
before squaring.  The negative controls are sharp: a signed linear sum of
instanton amplitudes can vanish while
\(\mathfrak a^\dagger W\mathfrak a\) is positive; using unamputated source
coordinates computes \(a^\dagger Z^\dagger WZa\), not the physical cut; and
dropping \(W\) replaces the declared observable by a locally inclusive one.
\end{controlledapproximation}

\section{First Cluster Corrections Are Amplitude Data}
\label{sec:instanton-first-cluster-amplitude-correction}

Equation~\eqref{eq:instanton-hard-assembled-channel} is a \(Q=1\) source
coefficient.  Taken alone, it leaves sector isolation from neighboring
topological sectors unproved.  The next obstruction lies in the connected
two-body amplitude in the same regulator, source frame, and physical
projection.

Let \(\sigma,\tau\in\{+,-\}\) denote instanton and anti-instanton charges,
and write \(z_1,z_2\) for their collective coordinates after the common
center has been separated.  In a retained two-body chart, let
\(\dd\nu_{\sigma\tau}^{\rm ord}(z_1,z_2)\) denote the ordered measure: the
first labelled entry has charge \(\sigma\), the second has charge \(\tau\),
and no pair symmetry factor is hidden in the measure.  The source-channel
cluster coefficient has the form
\begin{equation}
  \mathcal A_{\sigma\tau}^{\rm conn}(\mathcal J)
  =
  \ee^{\ii(\sigma+\tau)\bar\theta}
  \int_{\mathcal W_{\sigma\tau}}
  \left[
    K_{\sigma\tau}^{\rm full}(z_1,z_2;\mathcal J)
    -
    K_{\sigma}^{(1)}(z_1;\mathcal J)
    K_{\tau}^{(1)}(z_2;\mathcal J)
  \right]\,
  \dd\nu_{\sigma\tau}^{\rm ord}(z_1,z_2).
  \label{eq:instanton-pair-connected-source-coefficient}
\end{equation}
The full two-body kernel includes the relative quasi-zero-mode interaction,
the zero-mode overlap matrix, the nonzero-mode determinant in the pair
background, the source insertion, and the final projection.  The second term
is the disconnected product of the two already-counted one-body channels.
Removing it is part of the amplitude calculation, not an aesthetic
normalization choice.  Without this subtraction the cluster coefficient
double counts the Poisson term.

For \(\sigma+\tau=0\), the pair is theta-neutral.  That does not make it
irrelevant.  An instanton--anti-instanton molecule may change a source
correlator through zero-mode overlap singular values or through the
source-dependent nonzero-mode quotient even though it contributes no
\(\bar\theta\)-curvature at this order.  For \(\sigma+\tau=\pm2\), the
same-charge pair carries a second theta harmonic; in a vacuum observable map
it is the source of the first \(\cos(2\bar\theta)\) correction described in
Controlled Approximation~\ref{ca:first-cluster-correction-dilute-instantons}.
Thus theta curvature, hard source coefficients, and neutral-pair source
correlators again read different pieces of the same finite-regulator
instanton ensemble.

\begin{controlledapproximation}[First cluster correction to a source amplitude]
\label{ca:instanton-first-cluster-amplitude-correction}
Fix a one-instanton source channel \(\mathcal C\) and its anti-instanton
conjugate in the same scheme.  A first cluster approximation to the selected
source amplitude has the ordered form
\begin{equation}
  \mathcal A_{\rm cl}^{(2)}(\mathcal J)
  =
  \mathcal A_+^{(1)}(\mathcal J)
  +\mathcal A_-^{(1)}(\mathcal J)
  +\frac12\sum_{\sigma,\tau\in\{+,-\}}
    \mathcal A_{\sigma\tau}^{\rm conn}(\mathcal J)
  +R_{\ge3}^{\rm cl}(\mathcal J).
  \label{eq:instanton-source-first-cluster-assembly}
\end{equation}
The displayed factor \(1/2\) is the Mayer pair symmetry factor for the
ordered species sum.  For \(\sigma=\tau\) it is the identical-pair \(1/2!\).
For \(\sigma\ne\tau\), the two ordered terms \(+-\) and \(-+\) together give
one unordered mixed pair when charge conjugation or the chosen source frame
identifies their coefficients.  Equivalently one may use unordered pair
measures and absorb these factors into the definitions of
\(\mathcal A_{\sigma\tau}^{\rm conn}\); that is a different convention and
must be stated before comparing amplitudes.
Let \(C_{\sigma\tau}^{0}(z_1,z_2;\mathcal J)\) be the retained leading
connected pair kernel after the disconnected one-body product in
\eqref{eq:instanton-pair-connected-source-coefficient} has been subtracted,
and set
\[
  H_{\sigma\tau}^{0}(\mathcal J)
  =
  \int_{\mathcal W_{\sigma\tau}}
  C_{\sigma\tau}^{0}(z_1,z_2;\mathcal J)\,
  \dd\nu_{\sigma\tau}^{\rm ord},
  \qquad
  M_{\sigma\tau}^{0}(\mathcal J)
  =
  \int_{\mathcal W_{\sigma\tau}}
  |C_{\sigma\tau}^{0}(z_1,z_2;\mathcal J)|\,
  \dd|\nu_{\sigma\tau}^{\rm ord}|.
\]
Assume the retained pair kernel differs from \(C_{\sigma\tau}^{0}\) by
multiplicative corrections
\[
  \mathcal R_y=1+\delta_y,\qquad |\delta_y|\le e_y,
  \qquad
  y\in\{{\rm qzm},{\rm ov},{\rm nz},{\rm src},{\rm phys}\},
\]
where the entries are, respectively, the relative quasi-zero-mode
interaction, zero-mode overlap determinant, nonzero-mode determinant mismatch,
source-kernel correction, and physical-projection mismatch.  Let
\[
  \epsilon_{\sigma\tau}^{\rm pair}
  =
  \prod_y(1+e_y)-1,
\]
and let \(R_{\sigma\tau}^{\rm tail}\) denote the absolute remainder from
relative-coordinate tails, collision faces, and the boundary of the two-body
window.  Then
\begin{equation}
  \left|
  \mathcal A_{\sigma\tau}^{\rm conn}(\mathcal J)
  -
  \ee^{\ii(\sigma+\tau)\bar\theta}
  H_{\sigma\tau}^{0}(\mathcal J)
  \right|
  \le
  \epsilon_{\sigma\tau}^{\rm pair}
  M_{\sigma\tau}^{0}(\mathcal J)
  +R_{\sigma\tau}^{\rm tail}.
  \label{eq:instanton-source-first-cluster-bound}
\end{equation}
The relative-coordinate finiteness needed to make
\(R_{\sigma\tau}^{\rm tail}\) small is the integrability condition of
Controlled Approximation~\ref{ca:instanton-two-body-cluster-integrability},
transported to the present source and projection data.

This assembly has three immediate negative controls.  First, replacing
\(K_{\sigma\tau}^{\rm full}\) by the product
\(K_{\sigma}^{(1)}K_{\tau}^{(1)}\) gives zero connected cluster data and only
recovers the Poisson approximation.  Second, using theta curvature to bound a
neutral \(I\bar I\) source coefficient misses the case
\(\sigma+\tau=0\), where the theta phase is trivial but the source kernel or
zero-mode overlap can be nonzero.  Third, a one-instanton sector-isolation
claim must budget the absolute size of the retained pair terms in
\eqref{eq:instanton-source-first-cluster-assembly}, including the displayed
symmetry factor; finite one-body density, by itself, does not control
\(R_{\rm sec}\).  Using the ordered pair sum without the \(1/2\) would double
count both identical pairs and the two ordered representatives of a mixed pair
whenever they describe one unordered physical molecule.
\end{controlledapproximation}

\begin{controlledapproximation}[Neutral-pair valley prescription]
\label{ca:instanton-neutral-pair-valley-prescription}
The neutral \(I\bar I\) channel has an additional prescription input.  In the
attractive relative quasi-zero-mode direction, a regulated valley coordinate
\(r\) can produce a lateral value rather than an ordinary positive molecule
integral.  For a selected source and projection coordinate, write the retained
neutral-pair functional in the form
\begin{equation}
  \mathcal V_{\pm}[\mathcal J]
  =
  \operatorname{PV}
  \int_{\mathcal W_{+-}}
  C_{+-}^{0}(r,u;\mathcal J)\,\dd r\,\dd u
  \ \pm\ \ii\pi\,\mathfrak r_{+-}(\mathcal J)
  +R_{\rm val,\pm}(\mathcal J),
  \label{eq:instanton-neutral-valley-lateral}
\end{equation}
where \(u\) denotes the remaining relative orientation, size, and source
variables, and \(\mathfrak r_{+-}(\mathcal J)\) is the residue or thimble
jump in that same channel.  The symbol \(\pm\) records the lateral
continuation used to define the attractive quasi-zero-mode integral.

The cancellation statement is meaningful only in the same observable
coordinate.  Suppose the perturbative sector for the same source and
projection has the lateral ambiguity
\begin{equation}
  \mathcal B_{\pm}[\mathcal J]
  =
  \mathcal B_{\rm PV}[\mathcal J]
  \mp\ii\pi\,\mathfrak r_{+-}(\mathcal J)
  +R_{\rm Bor,\pm}(\mathcal J).
  \label{eq:instanton-neutral-valley-borel-partner}
\end{equation}
Then the combined neutral contribution satisfies the exact identity
\begin{equation}
\begin{aligned}
  &\bigl(\mathcal B_{+}[\mathcal J]+\mathcal V_{+}[\mathcal J]\bigr)
  -
  \bigl(\mathcal B_{-}[\mathcal J]+\mathcal V_{-}[\mathcal J]\bigr) \\
  &\qquad =
  R_{\rm Bor,+}-R_{\rm Bor,-}+R_{\rm val,+}-R_{\rm val,-}.
\end{aligned}
  \label{eq:instanton-neutral-valley-ambiguity-cancellation}
\end{equation}
Thus the residue \(\mathfrak r_{+-}\) must carry the zero-mode overlap,
nonzero-mode determinant, source insertion, endpoint window, and physical
projection of the channel being claimed.  A vacuum valley residue transports
to a hard source, pole, or inclusive observable only after those maps have
been included.

If
\[
  |R_{\rm Bor,+}-R_{\rm Bor,-}|\le B_{\rm Bor},
  \qquad
  |R_{\rm val,+}-R_{\rm val,-}|\le B_{\rm val},
\]
then the residual prescription dependence of the combined neutral coordinate
is bounded by \(B_{\rm Bor}+B_{\rm val}\).  A pair-only lateral value, a
principal-value valley integral with the residue discarded, or a cancellation
checked in a different source frame gives a different physical assertion.
\end{controlledapproximation}

\section{Observable Maps Are Not Interchangeable}
\label{sec:instanton-observable-handoffs}

The assembled coefficient in
\eqref{eq:instanton-hard-assembled-channel} is still not a universal
``instanton contribution.''  It is a coefficient in a chosen source and
projection coordinate.  The same \(Q=\pm1\) saddle data can feed several
different QFT observables only after different final maps have been applied.
Those maps have different dimensions, positivity properties, and limiting
procedures.

\begin{controlledapproximation}[Observable maps from instanton channels]
\label{ca:instanton-observable-handoff-ledger}
Let \(\mathcal X_{+}\) be the finite-regulator \(Q=1\) one-instanton
amplitude density after the collective density, nonzero-mode determinant,
zero-mode saturation, source matching, and endpoint window have all been
assembled.  Let \(\mathcal X_{-}\) be the corresponding anti-instanton
density.  Four common observable maps are these.
\begin{enumerate}[label=(\roman*)]
\item A hard source coefficient extracts
\(\mathfrak L_{\rm hard}[\mathcal X_+]\) and still needs the Haar projector,
amputation convention, and spectral or OPE projection.
\item A theta curvature extracts the vacuum activity
\(\zeta_+\ee^{\ii\bar\theta}+\zeta_-\ee^{-\ii\bar\theta}\) and still needs
zero-mode saturation and a cluster window.
\item A \(U(1)_A\)-odd susceptibility extracts the zero-mode overlap singular
values \(s_\alpha\) from an \(I\bar I\) ensemble and still needs the ensemble
weight and the thermodynamic/chiral order of limits.
\item A real-time axial relaxation rate extracts \(\Gamma_{\rm CS}\) from the
retarded zero-frequency slope and still needs KMS continuation and the axial
susceptibility.
\end{enumerate}
In a CP-symmetric dilute vacuum window, \(\zeta_+=\zeta_-=\zeta\) gives
\[
  \frac{Z_{\rm dig}(\bar\theta)}{Z_0}
  =
  \exp\!\left(2V_4\zeta\cos\bar\theta\right),
  \qquad
  \chi_{\rm top}^{\rm dig}=2\zeta,
  \qquad
  b_2^{\rm dig}=-\frac1{12},
\]
as in
\eqref{eq:dilute-instanton-gas-partition}--\eqref{eq:dilute-instanton-gas-cumulants}.
This map is a vacuum partition-function statement.  If one light Dirac
flavor has an unsaturated zero mode, the corresponding mass-saturated vacuum
activity is zero, even though a differentiated fermionic source coefficient
such as \eqref{eq:instanton-hard-assembled-channel} can be nonzero.

The \(U(1)_A\)-odd susceptibility map is different.  In an
instanton--anti-instanton ensemble with zero-mode overlap singular values
\(s_\alpha\), the zero-mode-zone contribution is governed by
\[
  \Delta_{\pi-\delta}^{\rm zm}(m)
  =
  4m^2\int_0^\infty
  \frac{\rho_{\rm zm}(s)}{(m^2+s^2)^2}\,\dd s
\]
after the exact-topology density has been removed in the thermodynamic
order of limits; see
\eqref{eq:instanton-zero-mode-zone-u1a-kernel}.  It is controlled by the
near-zero singular-value law, not by the one-body Poisson activity alone.
Likewise the real-time axial relaxation map is
\[
  \gamma_5^{\rm top}
  =
  \frac{(2N_f)^2\Gamma_{\rm CS}}{2\chi_5T},
\]
as in \eqref{eq:qcd-massless-topological-axial-relaxation}; it needs the
retarded Chern--Simons diffusion slope, not only the Euclidean theta
curvature.

Consequently a valid instanton calculation must name the observable functional.
A hard four-source coefficient is not a topological susceptibility; a
Euclidean topological susceptibility is not a real-time sphaleron or axial
relaxation rate; and a dilute instanton curvature is not the
Witten--Veneziano pure-Yang--Mills curvature unless a same-scheme curvature
comparison and the pseudoscalar pole-normalization residuals have also been
supplied.
\end{controlledapproximation}

The physics content is the non-interchangeability.  The BPST saddle fixes the
topological charge, the zero-mode chirality, and the local semiclassical
weights.  The observable decides which part of that data survives: a
differentiated source sees fermion zero-mode slots; a theta curvature sees a
vacuum activity with all zero modes saturated; a \(U(1)_A\) susceptibility sees
the near-zero spectrum of an ensemble; and a real-time rate sees a retarded
low-frequency slope.

\section{Finite-Cell Control Model}
\label{sec:finite-cell-instanton-channel-control}

A continuum proof of an instanton amplitude is analytic.  Nevertheless, the
logic which prevents a moduli-only shortcut is already visible in a finite
cell model.  Let the retained instanton window be decomposed into finitely
many cells \(i\).  Let
\[
  c_i
  =
  w_i\,d_i\,\det B_i\,p_i
\]
denote the leading cell contribution after the positive collective weight
\(w_i\), determinant factor \(d_i\), zero-mode source determinant
\(\det B_i\), and physical projection weight \(p_i\) have all been inserted.

\begin{controlledapproximation}[Finite-cell instanton channel control]
\label{ca:finite-cell-instanton-channel-control}
In a finite retained-cell channel, suppose each leading cell contribution
\(c_i\) is replaced by \(c_i(1+\delta_i)\), with
\(|\delta_i|\le\epsilon_i\), and suppose the omitted projection, sector,
infrared, cut, and scheme remainders have absolute majorant \(R_{\rm ext}\).
Then
\begin{equation}
  \left|
  \mathcal A_{\rm phys}
  -\sum_i c_i
  \right|
  \le
  \sum_i |c_i|\,\epsilon_i+R_{\rm ext}.
  \label{eq:finite-cell-instanton-channel-bound}
\end{equation}
If the source matrix on a cell is perturbed by
\(B_i\mapsto B_i+E_i\), with \(B_i\) invertible and
\(\eta_i=\|B_i^{-1}E_i\|\) in an operator norm for which
\(|\operatorname{tr}X|\le2\|X\|\) and \(|\det X|\le\|X\|^2\) on
\(2\times2\) matrices, then
\begin{equation}
  \frac{|\det(B_i+E_i)-\det B_i|}{|\det B_i|}
  \le
  2\eta_i+\eta_i^2 .
  \label{eq:finite-cell-source-determinant-stability}
\end{equation}
The first estimate is the triangle inequality applied after the channel has
already been projected to the same observable coordinate.  It would be false
as a relative statement without a noncancellation margin for
\(\sum_i c_i\).  For the determinant estimate, write
\[
  \det(B_i+E_i)=\det B_i\,
  \det(1+B_i^{-1}E_i).
\]
For a \(2\times2\) matrix \(X\),
\(\det(1+X)-1=\operatorname{tr}X+\det X\).  The stated norm bounds give
\eqref{eq:finite-cell-source-determinant-stability}.
\end{controlledapproximation}

The finite-cell model has two uses.  First, it gives a transparent negative
control: deleting \(\det B_i\), \(p_i\), or \(R_{\rm ext}\) changes the
physical statement, even if the collective-coordinate weights are unchanged.
Second, it records why the continuum estimates in
Chapter~\ref{ch:volume-ii-20} carry so many named residuals.  They
are not decorative error terms.  They are the places where a Euclidean saddle
coefficient becomes, or fails to become, a physical amplitude.
