\chapter{Local Data, Spectral Measures, and Green Functions}
\label{ch:volume-ii-local-data}

\section{Hilbert-Space Data and Poincare Generators}

The constructions in this volume use the local Hilbert-space data introduced
earlier. Fix a spacetime dimension \(D\), write \(d=D-1\), and use the
mostly-plus metric
\[
  \eta_{\mu\nu}=\operatorname{diag}(-,+,\ldots,+),
  \qquad
  p^2=\eta_{\mu\nu}p^\mu p^\nu .
\]
The data are:
\begin{enumerate}
  \item a Hilbert space \(\Hilb\);
  \item a strongly continuous unitary representation
  \(U(\Lambda,a)\) of the proper orthochronous Poincare group, or of its
  spin cover when spinor fields are present;
  \item commuting self-adjoint translation generators \(P^\mu\), normalized by
  \[
    U(\mathbf 1,a)=\exp\{\ii a^\mu P_\mu\};
  \]
  \item the spectrum condition
  \[
    \operatorname{Spec}(P)\subset\overline V_+
    =
    \{p\in\mathbb R^D:p^0\ge0,\ p^2\le0\};
  \]
  \item a Poincare-invariant unit vector \(\Omega\in\Hilb\);
  \item local observables \(\Obs(O)\) assigned to bounded spacetime regions
  \(O\), or equivalently a compatible family of operator-valued distributions
  \(\widehat\Phi_\alpha(x)\) on a common invariant dense domain.
\end{enumerate}
The group law is
\[
  U(\Lambda',a')U(\Lambda,a)
  =
  U(\Lambda'\Lambda,\Lambda'a+a') .
\]
For an infinitesimal transformation
\[
  \Lambda^\mu{}_\nu=\delta^\mu{}_\nu+\omega^\mu{}_\nu,
  \qquad
  a^\mu=\epsilon^\mu,
  \qquad
  \omega_{\mu\nu}=-\omega_{\nu\mu},
\]
the generators \(P^\mu\) and \(J^{\mu\nu}\) are defined by
\[
  U(\Lambda,a)
  =
  1+\ii\epsilon^\mu P_\mu
  +\frac{\ii}{2}\omega_{\mu\nu}J^{\mu\nu}
  +O(\epsilon^2,\omega^2,\epsilon\omega).
\]
When a conserved stress tensor \(T^{\mu\nu}\) is available, the corresponding
charges have the structural form
\[
  P^\mu=\int_{\Sigma_t}\dd^d\vec x\,T^{0\mu}(t,\vec x),
\]
and
\[
  J^{\mu\nu}
  =
  \int_{\Sigma_t}\dd^d\vec x\,
  \bigl(x^\mu T^{0\nu}(t,\vec x)-x^\nu T^{0\mu}(t,\vec x)\bigr).
\]
In an interacting continuum theory these formulae require the same regulator,
domain, and renormalization choices as the stress tensor itself. The displayed
integrals specify the covariance charges when those choices have been made.

The vacuum condition is
\[
  P^\mu\Omega=0,
  \qquad
  J^{\mu\nu}\Omega=0 .
\]
A local field with Lorentz index \(\alpha\) transforms as
\[
  U(\Lambda,a)^{-1}\widehat\Phi_\alpha(x)U(\Lambda,a)
  =
  R(\Lambda)_\alpha{}^\beta\,
  \widehat\Phi_\beta(\Lambda x+a),
\]
where \(R\) is the finite-dimensional Lorentz representation appropriate to
the field.  This active/passive convention is the one used in the scattering
chapters: the unitary representation translates states by \(U\), while the
coordinate argument of a field is translated by conjugation with \(U^{-1}\) on
the left. Since fields are distributions, this identity means equality after
smearing against test functions. For translations it gives
\[
  \widehat\Phi_\alpha(x)
  =
  \exp\{-\ii x^\mu P_\mu\}\widehat\Phi_\alpha(0)
  \exp\{\ii x^\mu P_\mu\}.
\]
Locality means that if \(O_1\) and \(O_2\) are spacelike separated, bounded
observables \(A_i\in\Obs(O_i)\) commute. For fields, the corresponding
statement is imposed after smearing with test functions whose supports are
spacelike separated.

These data support several constructions, each with its own hypotheses:
\[
\begin{tikzpicture}[>=Stealth,scale=0.92,
  box/.style={draw,thick,align=center,minimum width=3.05cm,
    minimum height=0.92cm,fill=blue!4},
  arr/.style={->,thick}]
  \node[box] (local) at (0,0) {Poincare-local\\Hilbert data};
  \node[box] (spectral) at (4.15,0) {spectral\\measures};
  \node[box] (greens) at (8.3,0) {Green-function\\distributions};
  \node[box,fill=green!5] (scatter) at (4.15,-1.85)
    {asymptotic states\\and kernels};
  \node[box,fill=orange!7] (ren) at (8.3,-1.85)
    {renormalized\\continuum limits};
  \draw[arr] (local)--(spectral);
  \draw[arr] (spectral)--(greens);
  \draw[arr] (spectral)--(scatter);
  \draw[arr] (greens)--(scatter);
  \draw[arr] (greens)--(ren);
\end{tikzpicture}
\]
The first arrow is the joint spectral theorem applied to \(P^\mu\). The arrows
to scattering use isolated one-particle spectral support and large-time
limits. The arrow to renormalized continuum limits uses regulated correlation
functions and locality of counterterms.

\section{One-Particle States Created by Local Fields}

Let \(m>0\) and let
\[
  \Sigma_m^+
  =
  \{p\in\mathbb R^D:p^0=\omega_{\vec p},\ p^2=-m^2\},
  \qquad
  \omega_{\vec p}=(\vec p^{\,2}+m^2)^{1/2}.
\]
An isolated one-particle species is represented by a spectral subspace
supported on \(\Sigma_m^+\) and separated from the rest of the spectrum by a
mass gap in a neighborhood of the shell. Delta-normalized momentum vectors
\(|\vec p\rangle_\delta\) obey
\[
  {}_\delta\langle\vec q\,|\vec p\rangle_\delta
  =
  \delta^d(\vec q-\vec p),
  \qquad
  P^i|\vec p\rangle_\delta=p^i|\vec p\rangle_\delta,
  \qquad
  P^0|\vec p\rangle_\delta=\omega_{\vec p}|\vec p\rangle_\delta .
\]
Covariantly normalized generalized vectors are
\[
  |\vec p\rangle
  =
  \bigl((2\pi)^d2\omega_{\vec p}\bigr)^{1/2}
  |\vec p\rangle_\delta,
  \qquad
  \langle\vec q\,|\vec p\rangle
  =
  (2\pi)^d2\omega_{\vec p}\,\delta^d(\vec q-\vec p).
\]

For a scalar local field \(\widehat\phi\), the one-particle matrix element is
specified by a nonnegative constant \(Z_\phi\):
\[
  \langle\vec p\,|\widehat\phi(0)|\Omega\rangle=Z_\phi^{1/2}.
\]
Equivalently, in delta normalization,
\[
  {}_\delta\langle\vec p\,|\widehat\phi(0)|\Omega\rangle
  =
  \left(\frac{Z_\phi}{(2\pi)^d2\omega_{\vec p}}\right)^{1/2}.
\]
Using translation covariance, the field applied to the vacuum has the
decomposition
\[
\begin{aligned}
  \widehat\phi(x)\Omega
  &=
  \int \dd^d\vec p\,
  \ee^{-\ii\vec p\cdot\vec x+\ii\omega_{\vec p}x^0}
  \left(\frac{Z_\phi}{(2\pi)^d2\omega_{\vec p}}\right)^{1/2}
  |\vec p\rangle_\delta
\\
  &\quad
  +\int_{\mathrm{cont}}\dd\alpha\,
  \ee^{-\ii p_\alpha\cdot x}
  \langle\alpha|\widehat\phi(0)|\Omega\rangle\,|\alpha\rangle .
\end{aligned}
\]
The second integral denotes the part of the spectral decomposition supported
away from the isolated one-particle shell, including multiparticle states
with the same quantum numbers as \(\widehat\phi\). Thus \(Z_\phi\) is the
one-particle overlap of the chosen interpolating field. If \(Z_\phi=0\), that
particular field does not create the particle from the vacuum, although a
different local or composite field may.

\section{Positive Spectral Measures}

Let \(E(\Delta)\) be the joint spectral projection of \(P^\mu\) for a Borel
set \(\Delta\subset\mathbb R^D\). If
\[
  v_\alpha=\widehat\Phi_\alpha(f)\Omega ,
\]
then
\[
  \nu_{\Phi,f}(\Delta)
  =
  \langle v_\alpha,E(\Delta)v_\alpha\rangle_{\Hilb}
\]
is a positive Borel measure on momentum space. It records the
energy-momentum content created from the vacuum by the smeared local field.

For a scalar field \(\widehat\phi\), translation invariance, Lorentz
covariance, the spectrum condition, and Hilbert-space positivity imply that
the two-point Wightman distribution has the K\"all\'en--Lehmann form
\[
  W_2(x-y)
  =
  \langle\Omega|\widehat\phi(x)\widehat\phi(y)|\Omega\rangle
  =
  \int_0^\infty \dd\rho_\phi(\mu^2)\,
  \Delta_+(x-y;\mu^2),
\]
where \(\dd\rho_\phi\) is a positive Borel measure on \([0,\infty)\), and
\[
  \Delta_+(x;\mu^2)
  =
  \int\frac{\dd^D k}{(2\pi)^{D-1}}\,
  \theta(k^0)\delta(k^2+\mu^2)\,\ee^{\ii k\cdot x}.
\]
Positivity of \(\dd\rho_\phi\) is not an additional dynamical assumption in
this derivation. It is the positivity of Hilbert-space norms applied to the
spectral projections of \(\widehat\phi(f)\Omega\).

If a stable scalar particle of mass \(m\) is created by \(\widehat\phi\), the
spectral measure contains an atom
\[
  \dd\rho_\phi(\mu^2)
  =
  Z_\phi\,\delta(\mu^2-m^2)\,\dd\mu^2
  +
  \dd\rho_{\phi,\mathrm c}(\mu^2),
  \qquad
  Z_\phi>0.
\]
The continuous part begins at the lowest invariant mass accessible to the
field with the same quantum numbers. If the first such continuum consists of
two particles of mass \(m\), the threshold is at \(\mu=2m\).

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.98]
  \draw[->,line width=0.7pt] (0,0)--(7.25,0) node[right] {$\mu$};
  \draw[->,line width=0.7pt] (0,0)--(0,3.1)
    node[above] {$\rho_\phi(\mu^2)$};
  \draw[blue!75!black,line width=1.35pt] (1.85,0)--(1.85,2.78);
  \node[blue!75!black,below] at (1.85,-0.1) {$m$};
  \node[blue!75!black,anchor=west] at (2.08,2.78)
    {$Z_\phi\delta(\mu^2-m^2)$};
  \draw[dashed,line width=0.75pt] (3.65,0)--(3.65,1.05);
  \node[red!75!black,below] at (3.65,-0.1) {$2m$};
  \draw[red!75!black,line width=1.4pt]
    (3.65,0.06) .. controls (3.95,0.85) and (4.45,1.42) .. (5.18,1.33)
    .. controls (5.9,1.24) and (6.38,0.78) .. (6.86,0.58);
  \node[red!75!black,align=center] at (5.78,2.24)
    {multiparticle\\continuum};
  \draw[red!75!black,->,line width=0.7pt] (5.56,1.78)--(5.08,1.24);
  \draw[decorate,decoration={brace,amplitude=5pt},gray!75]
    (1.98,0.55)--(3.45,0.55);
  \node[gray!70!black] at (2.72,0.92) {gap};
\end{tikzpicture}
\end{center}

The atom supplies the one-particle Hilbert space and the pole residue used in
LSZ. The continuous support supplies threshold data, branch cuts of Green
functions, and the channels in which bound states or resonances may appear.

\section{Lorentzian and Euclidean Green Functions}

For local fields \(\widehat\Phi_{a_1},\ldots,\widehat\Phi_{a_N}\), where
\(a_i\) denotes all tensor, spinor, and species labels, define the
time-ordered distribution
\[
  G_{a_1\cdots a_N}(x_1,\ldots,x_N)
  =
  \langle\Omega|
  \operatorname T\{
    \widehat\Phi_{a_1}(x_1)\cdots
    \widehat\Phi_{a_N}(x_N)
  \}
  |\Omega\rangle .
\]
This is a distributional matrix element. Products at coincident points require
a specified time-ordering prescription or a renormalized definition of
composite insertions.

For a scalar field, the same positive spectral measure gives
\[
  G_2(x)
  =
  \int_0^\infty \dd\rho_\phi(\mu^2)\,\Delta_F(x;\mu^2),
\]
with
\[
  \Delta_F(x;\mu^2)
  =
  -\ii
  \int\frac{\dd^D k}{(2\pi)^D}\,
  \frac{\ee^{\ii k\cdot x}}{k^2+\mu^2-\ii0}.
\]
The \(-\ii0\) is the boundary-value prescription associated with time
ordering in the mostly-plus convention.

Euclidean points are written \(x_E=(\vec x,\tau)\in\mathbb R^D\), with
\(x^0=-\ii\tau\). When Schwinger functions \(S_N\) have been constructed and
satisfy the reconstruction hypotheses, the scalar two-point Schwinger
function has the representation
\[
  S_2(x_E)
  =
  \int_0^\infty \dd\rho_\phi(\mu^2)\,\Delta_E(x_E;\mu^2),
\]
where
\[
  \Delta_E(x_E;\mu^2)
  =
  \int\frac{\dd^D k_E}{(2\pi)^D}\,
  \frac{\ee^{\ii k_E\cdot x_E}}{k_E^2+\mu^2}.
\]
The passage between Wightman, time-ordered, and Euclidean functions is
controlled by spectral support, analyticity domains, and reconstruction
conditions. In perturbation theory it is implemented as a contour and
boundary-value calculation. In a constructive Euclidean framework it is part
of the reconstruction problem.

\section{Wick Rotation and Regulated Functional Integrals}

For ordered insertions in an analyticity domain, the rotation from Lorentzian
time to Euclidean time is represented by
\[
  x_a^0=\ee^{-\ii\alpha}\tau_a,
  \qquad
  \alpha:0\longrightarrow \frac{\pi}{2},
  \qquad
  x_a^D=\tau_a .
\]
At \(\alpha=\pi/2\), the insertions lie on the Euclidean time axis.

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.86]
  \draw[->,line width=0.7pt] (-4.2,0)--(4.4,0)
    node[below right] {$\operatorname{Re}x^0$};
  \draw[->,line width=0.7pt] (0,-3.4)--(0,3.0)
    node[above] {$-\operatorname{Im}x^0$};
  \draw[green!45!black,line width=1.25pt,->]
    (-3.6,-1.05) .. controls (-2.05,-0.55) and (-0.7,-0.1) .. (0,0);
  \draw[green!45!black,line width=1.25pt]
    (0,0) .. controls (1.15,0.24) and (2.45,0.53) .. (3.45,0.78);
  \node[green!45!black,below left] at (-3.5,-1.0) {$C$};
  \foreach \x/\y/\lab in {-2.9/-0.86/x_n^0,1.12/0.25/x_2^0,2.35/0.54/x_1^0} {
    \fill[red!75!black] (\x,\y) circle (2.2pt);
    \node[red!75!black,below] at (\x,\y-0.06) {\(\lab\)};
  }
  \node[red!75!black] at (-1.85,-0.48) {\(\cdots\)};
  \draw[red!75!black,->,line width=0.75pt] (1.62,0.42)--(2.42,0.6);
  \node[red!75!black,above] at (2.22,0.88) {time ordered};
  \foreach \y/\lab in {2.0/\tau_1,0.82/\tau_2,-2.55/\tau_n} {
    \fill[purple!80!black] (0,\y) circle (2.2pt);
    \node[purple!80!black,right] at (0.15,\y) {\(\lab\)};
  }
  \node[purple!80!black,right] at (0.16,-0.92) {\(\vdots\)};
  \node[purple!80!black,below right] at (0.18,-3.0) {Euclidean};
  \draw[orange!85!black,line width=1.0pt,->]
    (2.18,0.76) .. controls (2.76,1.52) and (1.15,1.92) .. (0.18,2.02);
  \draw[orange!85!black,line width=1.0pt,->]
    (-2.7,-1.03) .. controls (-2.0,-2.2) and (-0.82,-2.52) .. (-0.15,-2.55);
  \node[blue!70!black,align=left] at (5.95,1.7)
    {\(\displaystyle x_a^0=\ee^{-\ii\alpha}\tau_a\)\\[3pt]
     \(\displaystyle \alpha:0\to\frac{\pi}{2}\)\\[3pt]
     \(\displaystyle x_a^D=\tau_a\)};
\end{tikzpicture}
\end{center}

For a scalar Lagrangian whose analytic continuation is defined by the chosen
regulator, write
\[
  S=\int\dd^D x\,\mathcal L[\phi],
  \qquad
  S_E=\int\dd^D x_E\,\mathcal L_E[\phi],
\]
with
\[
  \mathcal L_E[\phi]
  =
  -\mathcal L[\phi]\big|_{x^0\to-\ii x_E^D,\ \vec x=\vec x_E}
\]
after the continuation of time derivatives and tensor components has been
specified. The Euclidean functional integral is then a regulated construction
of Schwinger functions. Let \(\Lambda\) be a regulator scale. A regulated
Euclidean field configuration space is denoted by \(\mathcal C_\Lambda\), and
its integration symbol by \(D_\Lambda\varphi\). This symbol denotes the
finite-dimensional or otherwise regulated integration supplied by the
regulator; it is not an unregulated measure on all distributions.

The regulated action is a functional
\[
  S_{\Lambda,E}[\varphi;g_I(\Lambda)]
  =
  \sum_I g_I(\Lambda) O_{I,\Lambda}[\varphi],
\]
where \(O_{I,\Lambda}\) are regulated local or quasi-local functionals and
\(g_I(\Lambda)\) are cutoff-dependent coefficients. For a source \(J\), the
regulated generating functional is
\[
  Z_\Lambda[J]
  =
  \int_{\mathcal C_\Lambda}D_\Lambda\varphi\,
  \exp\left(
    -S_{\Lambda,E}[\varphi;g_I(\Lambda)]
    +\int \dd^D x_E\,J(x_E)\varphi(x_E)
  \right).
\]
Regulated \(N\)-point functions are obtained by differentiating
\[
  \langle\varphi(x_1)\cdots\varphi(x_N)\rangle_\Lambda
  =
  \frac{1}{Z_\Lambda[0]}
  \frac{\delta^N Z_\Lambda[J]}
  {\delta J(x_1)\cdots\delta J(x_N)}
  \bigg|_{J=0}.
\]

A continuum Schwinger function is a limit
\[
  S_N(x_1,\ldots,x_N)
  =
  \lim_{\Lambda\to\infty}
  \langle\varphi(x_1)\cdots\varphi(x_N)\rangle_\Lambda
\]
in a specified distribution topology, after choosing the cutoff dependence of
the coefficients \(g_I(\Lambda)\). Existence of this limit is a dynamical
claim. Perturbative renormalization proves the existence of finite formal
coefficients order by order in selected classes of theories; constructive
methods prove stronger nonperturbative statements in selected models.

\begin{center}
\begin{tikzpicture}[>=Stealth,scale=0.94,
  box/.style={draw,thick,align=center,minimum width=3.25cm,
    minimum height=0.86cm,fill=green!5},
  arr/.style={->,thick}]
  \node[box] (cutoff) at (0,0)
    {cutoff data\\\(\mathcal C_\Lambda,S_{\Lambda,E}\)};
  \node[box] (coeff) at (4.35,0)
    {local coefficients\\\(g_I(\Lambda)\)};
  \node[box] (corr) at (8.7,0)
    {regulated\\correlators};
  \node[box,fill=orange!7] (limit) at (4.35,-1.85)
    {distributional\\continuum limit};
  \draw[arr] (cutoff)--(coeff);
  \draw[arr] (coeff)--(corr);
  \draw[arr] (corr)--(limit);
  \draw[arr] (coeff)--(limit);
\end{tikzpicture}
\end{center}
