\chapter{Chiral Axial Anomalies}
\label{ch:volume-ii-20}
\ConventionsLorentzian


\section{Classical Chiral Currents}

Let \(\psi\) be a Dirac fermion in four dimensions.  Throughout this chapter
the Lorentzian metric is mostly plus and the Dirac adjoint is
\[
  \bar\psi:=\ii\psi^\dagger\gamma^0 .
\]
These are the gamma-matrix conventions developed in
Chapter~\ref{ch:spinor-fields-fermionic-path-integrals} and centralized in
Section~\ref{sec:spinor-conventions}.  The anomaly signs below use the same
orientation, chirality matrix, and dimensional-continuation prescription.
The flat oriented coordinate systems used in the perturbative calculations are
fixed by
\[
  \epsilon^{01}=+1
  \quad\hbox{in two dimensions},
  \qquad
  \epsilon^{0123}=+1
  \quad\hbox{in four dimensions}.
\]
The symbols \(\epsilon^{\mu\nu}\) and \(\epsilon^{\mu\nu\rho\sigma}\) are
totally antisymmetric with these normalizations.  In four dimensions the
chirality convention is
\[
  \gamma_5=-\ii\gamma^0\gamma^1\gamma^2\gamma^3,
  \qquad
  \operatorname{tr}
  \bigl(\gamma_5\gamma^\mu\gamma^\nu\gamma^\rho\gamma^\sigma\bigr)
  =
  4\ii\,\epsilon^{\mu\nu\rho\sigma}.
\]
Gauge generators in this chapter use the same Hermitian convention as
Chapter~\ref{ch:vol2-classical-yang-mills-matter}: if
\(A_\mu=A_\mu^a t^a\), then
\[
  [t^a,t^b]=\ii f^c{}_{ab}t^c,
  \qquad
  \operatorname{tr}(t^a t^b)=\delta^{ab}
\]
unless a different trace is explicitly declared.  The trace-form anomaly
expressions are written in terms of the matrix-valued curvature
\[
  F_{\mu\nu}
  =
  \partial_\mu A_\nu-\partial_\nu A_\mu-\ii[A_\mu,A_\nu],
\]
so a change of generator normalization rescales the component fields and
coupling coordinates but not the matrix trace formula itself.
With this adjoint the Hermitian bilinear current associated with a real
phase rotation contains an explicit factor of \(\ii\).  Couple \(\psi\) to a
background or dynamical gauge field through
\[
  D_\mu\psi=\partial_\mu\psi-\ii A_\mu\psi .
\]

\begin{definition}[Anomaly representative conventions]
\label{def:anomaly-representative-conventions}
An anomaly formula in this chapter means a renormalized local Ward identity
for specified composite currents after four pieces of data have been fixed.
First, the gauge and vector Ward identities that define the background
coupling are preserved by the regulator and by the counterterm convention.
Second, the remaining local term is represented in the Hermitian trace
normalization stated above, namely
\(\operatorname{tr}(t^at^b)=\delta^{ab}\) unless another trace is declared.
Third, the orientation and chirality matrix are the displayed ones.  Fourth,
local counterterms have been chosen; changing them changes the representative
of the anomaly but not its BRST cohomology class.

The perturbative triangle calculations below use dimensional reduction in
the following restricted sense: the loop momentum is analytically continued
while \(\gamma_5\) and the antisymmetric tensor are kept in the physical
two- or four-dimensional subspace.  The heat-kernel calculation in the
measure section is instead a spectral regulator for a Berezinian trace.  No
Pauli--Villars auxiliary field is used, and no claim is made that
dimensional continuation by itself defines a nonperturbative fermionic path
integral.
\end{definition}

\begin{warning}[Anomaly signs are tied to the full convention datum]
The anomaly coefficients in this chapter are not portable by changing one
symbol at a time.  The signs and factors use simultaneously the mostly-plus
Dirac adjoint, \(\gamma_5=-\ii\gamma^0\gamma^1\gamma^2\gamma^3\),
\(\epsilon^{0123}=+1\), the Hermitian covariant derivative
\(D_\mu=\partial_\mu-\ii A_\mu\), the curvature
\(F_{\mu\nu}=\partial_\mu A_\nu-\partial_\nu A_\mu-\ii[A_\mu,A_\nu]\), and
the trace convention fixed in Chapter~\ref{ch:vol2-classical-yang-mills-matter}.
Changing any one of these choices requires translating the whole convention
datum, including the topological charge normalization and the local
counterterm representative.  Dimensional reduction below is only the
perturbative regulator used to compute the displayed Ward-identity
representative; it is not being used as a nonperturbative definition of the
fermionic path integral.
\end{warning}

\begin{proposition}[Anomaly-coordinate comparison]
\label{prop:anomaly-coordinate-comparison}
Fix a regulated background-field groupoid
\(\mathcal B_\Lambda//\mathcal G_\Lambda\), with right action
\(A\mapsto A^g\), and a local regulator chart in which the chiral fermion
functional is a scalar \(W_\Lambda[A]\).  Then the finite cochain
\[
  C_\Lambda(g;A)=W_\Lambda[A^g]-W_\Lambda[A]
\]
is the additive coordinate of the anomaly line.  More precisely:
\begin{enumerate}
  \item functoriality of the groupoid action gives
  \(C_\Lambda(gh;A)=C_\Lambda(h;A^g)+C_\Lambda(g;A)\);
  \item a local counterterm \(B_\Lambda[A]\) changes this coordinate by the
  coboundary \(B_\Lambda[A^g]-B_\Lambda[A]\);
  \item after choosing a local frame \(s_A\) for the anomaly line, the
  exponentiated transport cocycle is
  \(\alpha_\Lambda(A,g)=\exp C_\Lambda(g;A)\), and the same counterterm is
  the inverse frame change \(s_A\mapsto \exp(-B_\Lambda[A])s_A\);
  \item the inflow bulk carries the inverse line coordinate
  \(\alpha_\Lambda(A,g)^{-1}\);
  \item differentiating \(C_\Lambda(\exp(t\zeta);A)\) at \(t=0\), then taking
  the local continuum representative, gives the consistent local anomaly
  density \(a_4^{(1)}(\zeta,A)\), whose relative BRST class is the
  perturbative obstruction.
\end{enumerate}
Consequently local BRST cancellation removes the obstruction on the identity
component of the background groupoid, while flat determinant- or Pfaffian-line
holonomies on based loops remain separate global anomaly data.
\end{proposition}

\begin{proof}
The first identity is just the scalar form of functoriality:
\[
  W_\Lambda[A^{gh}]-W_\Lambda[A]
  =
  \bigl(W_\Lambda[(A^g)^h]-W_\Lambda[A^g]\bigr)
  +
  \bigl(W_\Lambda[A^g]-W_\Lambda[A]\bigr).
\]
Replacing \(W_\Lambda\) by \(W_\Lambda+B_\Lambda\) gives the displayed
coboundary.  If the line is locally framed by \(s_A\) and
\[
  T_gs_A=\alpha_\Lambda(A,g)s_{A^g},
\]
then the frame \(s'_A=\exp(-B_\Lambda[A])s_A\) gives
\[
  T_gs'_A
  =
  \exp(-B_\Lambda[A])\alpha_\Lambda(A,g)
  \exp(B_\Lambda[A^g])s'_{A^g},
\]
which is the exponential of the same additive coboundary.  The bulk inflow
line is the inverse anomaly line, so its transport cocycle is
\(\alpha_\Lambda^{-1}\), and tensoring boundary and bulk transports gives the
identity transport.

For a one-parameter gauge transformation \(g(t)=\exp(t\zeta)\), the derivative
of the finite cocycle is
\[
  \frac{d}{dt}\bigg|_{t=0}C_\Lambda(g(t);A)
  =
  \delta_\zeta W_\Lambda[A].
\]
After subtracting local counterterms and passing to the continuum local
functional, this derivative is represented by
\(\int_Ma_4^{(1)}(\zeta,A)\).  Differentiating the finite cocycle identity in
two infinitesimal directions gives the Wess--Zumino consistency condition,
and replacing the even gauge parameters by the odd ghost gives the relative
BRST cocycle in \(H^{1,4}(s\mid d)\), developed later in
Proposition~\ref{prop:local-anomalies-relative-brst-classes}.  A local
counterterm changes the representative by the corresponding relative
coboundary.  This infinitesimal argument only sees the identity component of
the background groupoid.  A flat but nontrivial line holonomy along a based
loop has zero local curvature and zero local descent class, but still
prevents a gauge-invariant global scalar section; that global obstruction is
the determinant/Pfaffian-line story developed in
Chapter~\ref{ch:curved-eta-invariants-global-anomalies}.
\end{proof}

For a mass \(m\), the Dirac Lagrangian is
\[
  \mathcal L
  =
  -\bar\psi\gamma^\mu D_\mu\psi
  -m\bar\psi\psi .
\]
When \(m=0\), the classical equations of motion imply conservation of the
vector and axial currents
\[
  j^\mu=\ii\bar\psi\gamma^\mu\psi,
  \qquad
  j_5^\mu=\ii\bar\psi\gamma^\mu\gamma_5\psi,
  \qquad
  \partial_\mu j^\mu=0,
  \qquad
  \partial_\mu j_5^\mu=0 .
\]
For \(m\neq0\), the axial divergence becomes
\[
  \partial_\mu j_5^\mu=2\ii m\,\bar\psi\gamma_5\psi
\]
with the \(\gamma_5\) convention stated above.
Here \(D_\mu\bar\psi=\partial_\mu\bar\psi+\ii\bar\psi A_\mu\), and the
classical equations of motion are
\((\gamma^\mu D_\mu+m)\psi=0\) and
\((D_\mu\bar\psi)\gamma^\mu-m\bar\psi=0\).  Since
\(\{\gamma_5,\gamma^\mu\}=0\), these equations give
\(\partial_\mu(\bar\psi\gamma^\mu\gamma_5\psi)=2m\bar\psi\gamma_5\psi\);
the explicit factor of \(\ii\) in \(j_5^\mu\) gives the displayed
normalization.

The corresponding infinitesimal transformations are
\[
  \delta_V\psi=\ii\alpha\psi,\qquad
  \delta_A\psi=\ii\beta\gamma_5\psi,
\]
where \(\alpha\) and \(\beta\) are constant real parameters.  Classically, for
massless fermions, both transformations are symmetries of the local
Lagrangian.  Quantum mechanically, the vector transformation is preserved
when it is gauged consistently, while the axial current acquires a local
divergence determined by the gauge background.

\section{A Two-Dimensional Contact-Term Calculation}

The local origin of an axial anomaly is already visible in a free
two-dimensional Dirac fermion.  Use the Lorentzian action
\[
  S_2=-\int\dd^2x\,\bar\psi\gamma^\mu\partial_\mu\psi,
  \qquad
  \{\gamma^\mu,\gamma^\nu\}=2\eta^{\mu\nu},
  \qquad
  \bar\psi=\ii\psi^\dagger\gamma^0,
\]
and define
\[
  \gamma:=\gamma^0\gamma^1,
  \qquad
  \gamma^2=1,
  \qquad
  j^\mu=\ii\bar\psi\gamma^\mu\psi,
  \qquad
  j_A^\mu=\ii\bar\psi\gamma^\mu\gamma\psi .
\]
The classical massless action is invariant under
\(\psi\mapsto \ee^{\ii\alpha}\psi\) and under
\(\psi\mapsto\ee^{\ii\beta\gamma}\psi\).  The quantum question is the
renormalized Ward identity obeyed by time-ordered current products.  Consider
\[
  G_A^{\mu\nu}(x)
  :=
  \bra{\vac}
  \operatorname T\,j_A^\mu(x)j^\nu(0)
  \ket{\vac} .
\]
At one loop the relevant momentum-space graph is shown in
Figure~\ref{fig:two-dimensional-axial-vector-loop}; the figure fixes the loop
momentum routing and distinguishes the axial-current insertion from the vector
current insertion used in the Ward identity below.
\begin{figure}[H]
  \centering
  \begin{tikzpicture}[>=Stealth,scale=1.0,baseline={(0,-0.2)},
    fermion/.style={line width=0.75pt},
    current/.style={circle,fill=black,inner sep=1.7pt}]
    \coordinate (L) at (-1.45,0);
    \coordinate (R) at (1.45,0);
    \draw[fermion,->] (L) to[bend left=55] node[above] {\(p+k\)} (R);
    \draw[fermion,->] (R) to[bend left=55] node[below] {\(k\)} (L);
    \draw[fermion,->] (-2.6,0) -- (L);
    \draw[fermion,->] (R) -- (2.6,0);
    \node[current,label=below:\(\gamma^\mu\gamma\)] at (L) {};
    \node[current,label=below:\(\gamma^\nu\)] at (R) {};
    \node[left] at (-2.6,0) {\(p\)};
  \end{tikzpicture}
  \caption{The two-dimensional current two-point function.  The left insertion
  is the axial current and the right insertion is the vector current.}
  \label{fig:two-dimensional-axial-vector-loop}
\end{figure}
\[
  -\int\frac{\dd^2k}{(2\pi)^2}\,
  \frac{
    \operatorname{tr}\!\left[
      \ii\gamma^\nu(\not p+\not k)
      \ii\gamma^\mu\gamma\,\not k
    \right]
  }{(k^2-\ii\epsilon)((p+k)^2-\ii\epsilon)} .
\]
Contracting with \(p_\mu\), Feynman parametrizing, and shifting
\(k\mapsto k-xp\) gives
\begin{align}
  \int\dd^2x\,\ee^{-\ii p\cdot x}
  \partial_\mu G_A^{\mu\nu}(x)
  &=
  \ii\int_0^1\dd x
  \int\frac{\dd^2k}{(2\pi)^2}\,
  \frac{
    \operatorname{tr}(\gamma^\nu\not k\,\not p\,\gamma\,\not k)
    -x(1-x)p^2
    \operatorname{tr}(\gamma^\nu\gamma\not p)
  }{(k^2+p^2x(1-x)-\ii\epsilon)^2}.
  \label{eq:two-dimensional-anomaly-shifted-integral}
\end{align}
The expression is a contact term, so its value depends on the regulator used
to define the composite current.  Continue the loop momentum to
\(D=2-\epsilon\) dimensions while keeping the chirality matrix
\(\gamma=\gamma^0\gamma^1\) two-dimensional.  Rotational averaging gives
\[
  \operatorname{tr}(\gamma^\nu\not k\,\not p\,\gamma\,\not k)
  \longmapsto
  \frac{k^2}{D}
  \operatorname{tr}(\gamma^\nu\gamma_\rho\not p\,\gamma\,\gamma^\rho)
  =
  \frac{2-D}{D}k^2
  \operatorname{tr}(\gamma^\nu\not p\,\gamma).
\]
The factor \(2-D=\epsilon\) multiplies the logarithmic divergence of the loop
integral and leaves a finite local term.  With
\(\operatorname{tr}(\gamma^\nu\gamma\not p)=2\epsilon^{\mu\nu}p_\mu\), one
finds
\[
  \int\dd^2x\,\ee^{-\ii p\cdot x}
  \partial_\mu G_A^{\mu\nu}(x)
  =
  -\frac1{\pi}\epsilon^{\mu\nu}p_\mu .
\]
Equivalently,
\begin{equation}
  \partial_\mu
  \bra{\vac}
  \operatorname T\,j_A^\mu(x)j^\nu(y)
  \ket{\vac}
  =
  \frac{\ii}{\pi}\epsilon^{\mu\nu}
  \partial_\mu\delta^{(2)}(x-y),
  \label{eq:two-dimensional-axial-vector-contact-term}
\end{equation}
whereas the vector Ward identity is preserved:
\[
  \partial_\mu
  \bra{\vac}
  \operatorname T\,j^\mu(x)j^\nu(y)
  \ket{\vac}
  =0 .
\]

Now couple the vector current to a background \(U(1)\) gauge field,
\[
  D_\mu=\partial_\mu-\ii A_\mu,
  \qquad
  \Delta S=\int\dd^2x\,j^\mu A_\mu .
\]
Expanding time-ordered correlators in powers of \(A\), the first-order term in
\eqref{eq:two-dimensional-axial-vector-contact-term} gives the local operator
identity
\begin{equation}
  \partial_\mu j_A^\mu(x)
  =
  \frac1{2\pi}\epsilon^{\mu\nu}F_{\mu\nu}(x),
  \qquad
  F_{\mu\nu}:=\partial_\mu A_\nu-\partial_\nu A_\mu .
  \label{eq:two-dimensional-axial-anomaly}
\end{equation}
Thus the anomaly is already present before any interaction among the fermions:
it is the finite remnant of a regulated contact term in a composite-current
Ward identity.

\begin{proposition}[Two-dimensional axial contact term]
\label{prop:two-dimensional-axial-contact-term}
For the massless two-dimensional Dirac fermion with current conventions of
this section, choose the renormalized current representative whose vector
Ward identity is preserved.  Then the axial-vector current product obeys
\[
  \partial_\mu
  \bra{\vac}
  \operatorname T\,j_A^\mu(x)j^\nu(y)
  \ket{\vac}
  =
  \frac{\ii}{\pi}\epsilon^{\mu\nu}
  \partial_\mu\delta^{(2)}(x-y),
  \qquad
  \partial_\mu
  \bra{\vac}
  \operatorname T\,j^\mu(x)j^\nu(y)
  \ket{\vac}
  =
  0 .
\]
Coupling \(j^\mu\) to a background Abelian gauge field with
\(D_\mu=\partial_\mu-\ii A_\mu\) gives the local operator identity
\[
  \partial_\mu j_A^\mu
  =
  \frac1{2\pi}\epsilon^{\mu\nu}F_{\mu\nu}.
\]
\end{proposition}

\begin{proof}
The one-loop contraction of
\(p_\mu\langle j_A^\mu j^\nu\rangle\) reduces, after Feynman
parametrization, to
\eqref{eq:two-dimensional-anomaly-shifted-integral}.  In dimensional
reduction the term with two loop momenta contains the factor
\((2-D)/D\), because only the physical two-dimensional gamma matrices
anticommute with \(\gamma\).  This evanescent factor multiplies the logarithmic
ultraviolet divergence of the integral and leaves the finite polynomial
\(-\pi^{-1}\epsilon^{\mu\nu}p_\mu\).  Fourier inversion gives the displayed
contact term.  The vector-current representative is fixed by subtracting the
local contact term that would otherwise violate the vector Ward identity.

Expanding the background-field insertion
\(\exp(\ii\int j^\mu A_\mu)\) to first order and using the contact term gives
\[
  \partial_\mu j_A^\mu(x)
  =
  \frac1{\pi}\epsilon^{\mu\nu}\partial_\mu A_\nu(x)
  =
  \frac1{2\pi}\epsilon^{\mu\nu}F_{\mu\nu}(x),
\]
where the last equality uses antisymmetry of \(\epsilon^{\mu\nu}\).
\end{proof}

\section{Contour Charges and Anomalous Charge Transport}

The contact term has a geometric interpretation.  In Euclidean two-space, a
conserved current \(j^\mu\) defines a charge on a closed contour \(C\),
\[
  Q_C=\int_C j^\mu n_\mu\,\dd s ,
\]
where \(n_\mu\) is the outward unit normal to the contour.  If
\(\partial_\mu j^\mu=0\) away from operator insertions, the contour can be
deformed through regions with no insertions.
Figure~\ref{fig:conserved-contour-charge-deformation} shows the allowed
contour deformation and the point at which an attempted deformation would cross
an operator insertion, where the contact term enters the Ward identity.
\begin{figure}[H]
  \centering
  \begin{tikzpicture}[>=Stealth,scale=0.95,baseline={(0,-0.15)}]
    \draw[thick] (-2.2,-1.15) rectangle (2.2,1.15);
    \node at (-1.25,0.2) {\(\mathcal O_1\)};
    \node at (1.15,-0.1) {\(\mathcal O_2\)};
    \draw[blue!70!black,thick,->] (-1.25,0.2) ellipse (0.45 and 0.32);
    \draw[blue!70!black,thick,->] (1.15,-0.1) ellipse (0.45 and 0.32);
    \draw[purple!70!black,thick,->] (0,0) ellipse (1.95 and 0.88);
    \node[blue!70!black] at (-1.25,-0.55) {\(C_1\)};
    \node[blue!70!black] at (1.15,-0.75) {\(C_2\)};
    \node[purple!70!black] at (0,1.42) {\(C\)};
    \node at (0,-1.55) {\(\displaystyle \int_D\partial_\mu j^\mu\,\dd^2x=0\)};
  \end{tikzpicture}
  \caption{For a conserved current, the charge contour can be moved from small
  circles around operator insertions to a larger contour without changing the
  insertion.}
  \label{fig:conserved-contour-charge-deformation}
\end{figure}
If \(C\) encloses all insertions and can be moved to infinity, the ordinary
global transformation leaves the correlation function invariant.

For the anomalous axial current the small contour still gives the local action
of the axial rotation on an operator.  The difference between the charge on a
large contour and the sum of charges on small contours is the flux of the
anomaly through the interpolating domain \(D\), as displayed in
Figure~\ref{fig:anomalous-contour-charge-deformation}:
\begin{figure}[H]
  \centering
  \begin{tikzpicture}[>=Stealth,scale=0.95,baseline={(0,-0.15)}]
    \draw[thick] (-2.3,-1.25) rectangle (2.3,1.25);
    \draw[purple!70!black,thick,->] (0,0) ellipse (2.0 and 1.0);
    \draw[blue!70!black,thick,->] (-0.9,0.25) ellipse (0.38 and 0.28);
    \draw[blue!70!black,thick,->] (0.85,-0.25) ellipse (0.38 and 0.28);
    \node at (-0.9,0.25) {\(\mathcal O_1\)};
    \node at (0.85,-0.25) {\(\mathcal O_2\)};
    \node[purple!70!black] at (1.7,1.05) {\(C_\infty\)};
    \node[blue!70!black] at (-1.35,-0.15) {\(C_1\)};
    \node[blue!70!black] at (1.3,-0.65) {\(C_2\)};
    \node at (0,0.78) {\(D\)};
  \end{tikzpicture}
  \caption{For the axial current in a background field, deforming the contour
  across the domain \(D\) changes the charge by the integral of the anomalous
  divergence.}
  \label{fig:anomalous-contour-charge-deformation}
\end{figure}
Using \eqref{eq:two-dimensional-axial-anomaly},
\begin{equation}
  \Delta Q_A
  =
  Q_{A,C_\infty}-\sum_iQ_{A,C_i}
  =
  \int_D\dd^2x\,\partial_\mu j_A^\mu
  =
  \frac1{2\pi}\int_D\dd^2x\,
  \epsilon^{\mu\nu}F_{\mu\nu}.
  \label{eq:two-dimensional-anomalous-charge-shift}
\end{equation}
Therefore an axial rotation by \(\alpha\) multiplies the correlation function
by the background-dependent phase
\[
  \exp(\ii\alpha\Delta Q_A).
\]
Equivalently, in the background-field effective action the transformation
shifts the action by
\[
  \Delta S
  =
  \frac{\alpha}{2\pi}\int\dd^2x\,
  \epsilon^{\mu\nu}F_{\mu\nu}.
\]

\section{The Four-Dimensional Triangle Contact Term}

The four-dimensional axial anomaly is produced by an analogous regulated
contact term with one higher short-distance singularity.  For a massless
Dirac fermion, study
\[
  \bra{\vac}
  \operatorname T\,j_5^\mu(x)j^\nu(y)j^\rho(z)
  \ket{\vac} .
\]
There are two oriented triangle graphs:
\begin{figure}[H]
  \centering
  \resizebox{0.95\textwidth}{!}{%
  \begin{tikzpicture}[>=Stealth,scale=0.92,baseline={(0,-0.2)},
    fermion/.style={line width=0.75pt},
    current/.style={circle,fill=black,inner sep=1.7pt}]
    \coordinate (A) at (-1.15,-0.85);
    \coordinate (B) at (1.15,-0.85);
    \coordinate (C) at (0,1.05);
    \draw[fermion,->] (A) -- node[below] {\(\ell+p\)} (B);
    \draw[fermion,->] (B) -- node[right] {\(\ell-k\)} (C);
    \draw[fermion,->] (C) -- node[left] {\(\ell\)} (A);
    \draw[fermion,->] (-2.25,-0.85) -- (A);
    \draw[fermion,->] (B) -- (2.25,-0.85);
    \draw[fermion,->] (C) -- (0,2.05);
    \node[current,label=below:\(j_5^\mu\)] at (A) {};
    \node[current,label=below:\(j^\rho\)] at (B) {};
    \node[current,label=right:\(j^\nu\)] at (C) {};
    \node[left] at (-2.25,-0.85) {\(p\)};
    \node[below right] at (2.25,-0.85) {\(-p-k\)};
    \node[above] at (0,2.05) {\(k\)};
    \node at (2.85,0) {\(+\)};
    \begin{scope}[xshift=6.2cm]
      \coordinate (A2) at (-1.15,-0.85);
      \coordinate (B2) at (1.15,-0.85);
      \coordinate (C2) at (0,1.05);
      \draw[fermion,->] (B2) -- node[below] {\(\ell\)} (A2);
      \draw[fermion,->] (A2) -- node[left] {\(\ell+p\)} (C2);
      \draw[fermion,->] (C2) -- node[right] {\(\ell+p+k\)} (B2);
      \draw[fermion,->] (-2.25,-0.85) -- (A2);
      \draw[fermion,->] (B2) -- (2.25,-0.85);
      \draw[fermion,->] (C2) -- (0,2.05);
      \node[current,label=below:\(j_5^\mu\)] at (A2) {};
      \node[current,label=below:\(j^\rho\)] at (B2) {};
      \node[current,label=right:\(j^\nu\)] at (C2) {};
      \node[above left] at (-2.25,-0.85) {\(p\)};
      \node[right] at (2.25,-0.85) {\(-p-k\)};
      \node[above] at (0,2.05) {\(k\)};
    \end{scope}
  \end{tikzpicture}
  }
  \caption{The two orientations of the one-axial, two-vector triangle.  The
  regulator is chosen so that the vector-current Ward identities are
  preserved.}
  \label{fig:four-dimensional-axial-vector-vector-triangle}
\end{figure}
The divergence of the axial insertion is the Fourier transform of a local
contact term.  In the dimensional-reduction implementation of the perturbative
regulator, set \(D=4-\epsilon\) and split the loop momentum as
\[
  \ell=\ell_\parallel+\ell_\perp ,
\]
where \(\ell_\parallel\) lies in the physical four-dimensional subspace.
Keep \(\gamma_5=-\ii\gamma^{0123}\) four-dimensional, so that
\(\not\ell_\parallel\) anticommutes with \(\gamma_5\) and
\(\not\ell_\perp\) commutes with \(\gamma_5\).  In the first triangle one uses
\[
  \not p\,\gamma_5
  =
  \bigl[(\not\ell+\not p)\gamma_5+\gamma_5\not\ell\bigr]
  -2\not\ell_\perp\gamma_5 .
\]
The two bracketed terms are adjacent inverse-propagator numerators.  In the two
oriented triangle graphs they reduce one fermion propagator and become a pair of
vector-current contact terms with opposite routing.  Here is the cancellation
explicitly for the leading routing in Figure
\ref{fig:four-dimensional-axial-vector-vector-triangle}.  Set
\[
  D_0=\ell^2,\qquad
  D_k=(\ell-k)^2,\qquad
  D_p=(\ell+p)^2 .
\]
The bracketed part of the first orientation gives
\[
\begin{aligned}
  \mathcal C^{\nu\rho}_1
  =
  -\ii\int\frac{\dd^D\ell}{(2\pi)^D}
  \bigg[
  &
  \frac{
    \operatorname{tr}\!\left[
      \gamma_5\not\ell\gamma^\nu(\not\ell-\not k)\gamma^\rho
    \right]
  }{D_0D_k}
\\
  &+
  \frac{
    \operatorname{tr}\!\left[
      \gamma_5\gamma^\nu(\not\ell-\not k)\gamma^\rho(\not\ell+\not p)
    \right]
  }{D_kD_p}
  \bigg].
\end{aligned}
\]
The first term came from the adjacent factor
\((\not\ell+\not p)^2=D_p\), after cyclically moving the first
\((\not\ell+\not p)\) to the end of the trace; the second came from
\(\not\ell^2=D_0\).  The bracketed part of the opposite orientation is
\[
\begin{aligned}
  \mathcal C^{\nu\rho}_2
  =
  -\ii\int\frac{\dd^D\ell}{(2\pi)^D}
  \bigg[
  &
  \frac{
    \operatorname{tr}\!\left[
      \gamma_5\not\ell\gamma^\rho
      (\not\ell+\not p+\not k)\gamma^\nu
    \right]
  }{\ell^2(\ell+p+k)^2}
\\
  &+
  \frac{
    \operatorname{tr}\!\left[
      \gamma_5\gamma^\rho
      (\not\ell+\not p+\not k)\gamma^\nu(\not\ell+\not p)
    \right]
  }{(\ell+p+k)^2D_p}
  \bigg].
\end{aligned}
\]
The dimensional-reduction regulator is defined so that the reduced
two-propagator integrals are translation invariant; this is the perturbative
implementation of the vector Ward identities.  In the second term of
\(\mathcal C_2\), set
\(\ell'=\ell+p+k\) and then rename \(\ell'\) as \(\ell\).  It has the same
denominator \(D_0D_k\) as the first term of \(\mathcal C_1\), while
\[
\begin{aligned}
  &\operatorname{tr}\!\left[
    \gamma_5\not\ell\gamma^\nu(\not\ell-\not k)\gamma^\rho
  \right]
  =
  4\ii\epsilon^{\alpha\nu\beta\rho}\ell_\alpha(\ell-k)_\beta,
\\
  &\operatorname{tr}\!\left[
    \gamma_5\gamma^\rho\not\ell\gamma^\nu(\not\ell-\not k)
  \right]
  =
  4\ii\epsilon^{\rho\alpha\nu\beta}\ell_\alpha(\ell-k)_\beta
  =
  -4\ii\epsilon^{\alpha\nu\beta\rho}\ell_\alpha(\ell-k)_\beta .
\end{aligned}
\]
Thus these two terms cancel.  In the first term of \(\mathcal C_2\), set
\(\ell'=\ell+k\).  It has the same denominator \(D_kD_p\) as the second term
of \(\mathcal C_1\), and
\[
\begin{aligned}
  &\operatorname{tr}\!\left[
    \gamma_5\gamma^\nu(\not\ell-\not k)\gamma^\rho(\not\ell+\not p)
  \right]
  =
  4\ii\epsilon^{\nu\alpha\rho\beta}
  (\ell-k)_\alpha(\ell+p)_\beta,
\\
  &\operatorname{tr}\!\left[
    \gamma_5(\not\ell-\not k)\gamma^\rho(\not\ell+\not p)\gamma^\nu
  \right]
  =
  4\ii\epsilon^{\alpha\rho\beta\nu}
  (\ell-k)_\alpha(\ell+p)_\beta
\\
  &\hspace{3.6cm}
  =
  -4\ii\epsilon^{\nu\alpha\rho\beta}
  (\ell-k)_\alpha(\ell+p)_\beta .
\end{aligned}
\]
The bracketed contact terms therefore cancel pairwise once the vector Ward
identities are imposed by the regulator.  The entire axial contact term is
carried by the evanescent \(\ell_\perp\) contribution.  For the first
orientation this contribution is
\[
  2\ii
  \int\frac{\dd^D\ell}{(2\pi)^D}
  \frac{
    \operatorname{tr}\!\left[
      \gamma_5\not\ell_\perp\not\ell\gamma^\nu
      (\not\ell-\not k)\gamma^\rho(\not\ell+\not p)
    \right]
  }{D_0D_kD_p}.
\]
The coefficient of the local two-vector contact term is obtained by keeping
the part with two external momenta and no physical loop momentum in the
\(\gamma_5\)-trace.  Terms with a symmetric pair of physical loop momenta
contracted into the antisymmetric tensor vanish after rotational averaging, and
terms with an unpaired perpendicular gamma matrix or the wrong number of
physical gamma matrices have zero \(\gamma_5\)-trace.  Using
\[
  \not\ell_\perp\not\ell_\perp=\ell_\perp^2,
  \qquad
  \operatorname{tr}\!
  \left(\gamma_5\gamma^\nu\gamma^\alpha\gamma^\rho\gamma^\beta\right)
  =
  4\ii\epsilon^{\nu\alpha\rho\beta},
\]
one obtains
\[
\begin{aligned}
  \operatorname{tr}\!\left[
    \gamma_5\not\ell_\perp\not\ell\gamma^\nu
    (\not\ell-\not k)\gamma^\rho(\not\ell+\not p)
  \right]_{\rm local}
  &=
  -\ell_\perp^2 k_\alpha p_\beta
  \operatorname{tr}\!
  \left(\gamma_5\gamma^\nu\gamma^\alpha\gamma^\rho\gamma^\beta\right)
\\
  &=
  4\ii\,\ell_\perp^2
  \epsilon^{\nu\rho\alpha\beta}k_\alpha p_\beta .
\end{aligned}
\]
Thus the first orientation contributes
\[
  -8\,\epsilon^{\nu\rho\alpha\beta}k_\alpha p_\beta\,
  I(k,p),
  \qquad
  I(k,p):=
  \int\frac{\dd^D\ell}{(2\pi)^D}
  \frac{\ell_\perp^2}{D_0D_kD_p}.
\]
The opposite orientation gives the same local tensor after
\((\nu,k)\leftrightarrow(\rho,-p-k)\).

It remains to evaluate \(I(k,p)\).  Take the external momenta nonexceptional
during the ultraviolet computation; the contact coefficient extracted below is
independent of this infrared choice.  Feynman parametrization gives
\[
\begin{aligned}
  I(k,p)
  &=
  2\int_0^1\dd x\int_0^1\dd y\int_0^1\dd z\,
  \delta(1-x-y-z)
  \int\frac{\dd^D r}{(2\pi)^D}
  \frac{r_\perp^2}{(r^2+\Delta)^{3}},
\\
  r&=\ell-y k+z p,
\\
  \Delta
  &=
  y(1-y)k^2+z(1-z)p^2+2yz\,k\cdot p-\ii0 .
\end{aligned}
\]
The shift does not change the perpendicular component because \(p\) and \(k\)
lie in the physical four-dimensional subspace.  In the continued
\(D=4-\epsilon\) dimensional rotation-invariant integral,
\[
  \int\dd^D r\,r_\perp^2 f(r^2)
  =
  \frac{D-4}{D}
  \int\dd^D r\,r^2 f(r^2)
  =
  -\frac{\epsilon}{4-\epsilon}
  \int\dd^D r\,r^2 f(r^2).
\]
Therefore
\[
\begin{aligned}
  \int\frac{\dd^D r}{(2\pi)^D}
  \frac{r_\perp^2}{(r^2+\Delta)^3}
  &=
  \frac{D-4}{D}
  \int\frac{\dd^D r}{(2\pi)^D}
  \frac{r^2}{(r^2+\Delta)^3}
\\
  &=
  \frac{D-4}{4}
  \frac{\Gamma(2-D/2)}{(4\pi)^{D/2}}
  \Delta^{D/2-2}
\\
  &=
  -\frac1{32\pi^2}+O(\epsilon).
\end{aligned}
\]
Since
\[
  2\int_0^1\dd x\int_0^1\dd y\int_0^1\dd z\,
  \delta(1-x-y-z)=1
\]
and
\(\Delta^{-\epsilon/2}=1+O(\epsilon)\), the required scalar integral is
\[
  I(k,p)
  =
  -\frac1{32\pi^2}.
\]
With the sign convention of \eqref{eq:abelian-axial-anomaly}, this gives
\begin{equation}
  \partial_\mu
  \bra{\vac}
  \operatorname T\,j_5^\mu(x)j^\nu(y)j^\rho(0)
  \ket{\vac}
  =
  \frac1{2\pi^2}
  \epsilon^{\nu\rho\alpha\beta}
  \partial_\alpha\delta^{(4)}(x)
  \partial_\beta\delta^{(4)}(y).
  \label{eq:four-dimensional-triangle-contact-term}
\end{equation}
The local coefficient is independent of the long-distance state; it is fixed
by the regulated short-distance singularity of the current product.

Coupling the vector current to a background field,
\[
  \Delta S=\int\dd^4x\,A_\mu j^\mu,
\]
and expanding the time-ordered exponential to second order, the contact term
\eqref{eq:four-dimensional-triangle-contact-term} yields
\[
  \partial_\mu j_5^\mu
  =
  \frac1{16\pi^2}
  \epsilon^{\mu\nu\rho\sigma}F_{\mu\nu}F_{\rho\sigma}
\]
for a unit-charge Dirac fermion in the convention of
\eqref{eq:abelian-axial-anomaly}.  A fermion of electric charge \(q\) coupled
with electromagnetic coupling \(e\) replaces \(F\) by \(eqF\), giving the
factor \(e^2q^2\) in the next section.

\begin{proposition}[Four-dimensional axial contact term]
\label{prop:four-dimensional-axial-contact-term}
For a massless four-dimensional Dirac fermion, in the anomaly representative
of Definition~\ref{def:anomaly-representative-conventions}, the regulated
one-axial two-vector current product has local divergence
\[
  \partial_\mu
  \bra{\vac}
  \operatorname T\,j_5^\mu(x)j^\nu(y)j^\rho(0)
  \ket{\vac}
  =
  \frac1{2\pi^2}
  \epsilon^{\nu\rho\alpha\beta}
  \partial_\alpha\delta^{(4)}(x)
  \partial_\beta\delta^{(4)}(y).
\]
Consequently, after coupling the vector current to a background Abelian field
and preserving the vector Ward identity, a unit-charge Dirac fermion obeys
\[
  \partial_\mu j_5^\mu
  =
  \frac1{16\pi^2}
  \epsilon^{\mu\nu\rho\sigma}F_{\mu\nu}F_{\rho\sigma}
\]
in the massless theory.
\end{proposition}

\begin{proof}
Contracting the axial vertex in the two oriented triangles writes the
numerator as vector-current contact terms plus the evanescent
\(\ell_\perp\) term.  Translation invariance of the dimensionally reduced
two-propagator integrals is precisely the regulator statement that the vector
Ward identities are preserved; with this prescription the vector-contact
terms cancel pairwise between the two orientations.  The remaining local
piece is
\[
  -8\,\epsilon^{\nu\rho\alpha\beta}k_\alpha p_\beta I(k,p)
\]
from one orientation, with the opposite orientation contributing the
corresponding exchanged tensor.  The Feynman-parameter integral has the
regulator-independent local limit
\[
  I(k,p)
  =
  \int\frac{\dd^{4-\epsilon}\ell}{(2\pi)^{4-\epsilon}}
  \frac{\ell_\perp^2}{\ell^2(\ell-k)^2(\ell+p)^2}
  =
  -\frac1{32\pi^2},
\]
because the factor \(D-4\) from \(\ell_\perp^2\) cancels the logarithmic
ultraviolet pole.  Substituting this value and Fourier transforming gives
\eqref{eq:four-dimensional-triangle-contact-term}.

The background-field identity follows by expanding the time-ordered
exponential to second order.  The factor \(1/2\) in the exponential is
cancelled by the two identical vector insertions in the contact term, and
antisymmetrizing the two derivatives of \(A\) rewrites the result as
\(\epsilon^{\mu\nu\rho\sigma}F_{\mu\nu}F_{\rho\sigma}/(16\pi^2)\) in the
normalization of \eqref{eq:abelian-axial-anomaly}.
\end{proof}

\section{The Abelian Axial Anomaly}

For a Dirac fermion of electric charge \(q\) coupled to an electromagnetic
field with coupling \(e\), the renormalized axial current may be chosen so
that
\begin{equation}
  \partial_\mu j_5^\mu
  =
  2\ii m\,\bar\psi\gamma_5\psi
  +
  \frac{e^2 q^2}{16\pi^2}
  \epsilon^{\mu\nu\rho\sigma}
  F_{\mu\nu}F_{\rho\sigma}.
  \label{eq:abelian-axial-anomaly}
\end{equation}
The coefficient is fixed by the short-distance singularity of the triangle
diagram with one axial-current insertion and two vector-current insertions.
The vector Ward identity is imposed because the vector current is the current
coupled to the gauge field.  The remaining local term appears in the axial
Ward identity.

Equation \eqref{eq:abelian-axial-anomaly} is a statement about a
renormalized composite operator identity.  Each symbol in it has a prescribed
meaning: \(F_{\mu\nu}\) is the electromagnetic field strength,
\(\epsilon^{\mu\nu\rho\sigma}\) is the four-dimensional antisymmetric symbol
with the orientation fixed at the start of the chapter,
\(\ii m\bar\psi\gamma_5\psi\) is the pseudoscalar mass-breaking operator in the
mostly-plus adjoint convention, and \(j_5^\mu\) is the renormalized current
whose vector-current Ward identities have been preserved.

The factor relating this component formula to the heat-kernel trace below is
as follows.  Define
\[
  \widetilde F^{\mu\nu}
  =
  \frac12\epsilon^{\mu\nu\rho\sigma}F_{\rho\sigma},
  \qquad
  F=\frac12F_{\mu\nu}\dd x^\mu\wedge\dd x^\nu .
\]
Then
\[
  F_{\mu\nu}\widetilde F^{\mu\nu}
  =
  \frac12\epsilon^{\mu\nu\rho\sigma}F_{\mu\nu}F_{\rho\sigma},
  \qquad
  F\wedge F
  =
  \frac14\epsilon^{\mu\nu\rho\sigma}
  F_{\mu\nu}F_{\rho\sigma}\,\dd^4x .
\]
The local heat-kernel trace
\eqref{eq:heat-kernel-anomaly-density} gives the index density
\((32\pi^2)^{-1}\epsilon^{\mu\nu\rho\sigma}
\operatorname{tr}_{R}(F_{\mu\nu}F_{\rho\sigma})\).  The Dirac axial Ward
identity has twice this density because the chiral change of variables rotates
\(\psi\) and \(\bar\psi\) with the same sign, giving the prefactor \(2\ii\) in
\eqref{eq:fujikawa-regulated-trace}.  For a single charge-one Abelian Dirac
fermion, the representation trace is trivial and
\[
  2\cdot
  \frac{1}{32\pi^2}
  \epsilon^{\mu\nu\rho\sigma}F_{\mu\nu}F_{\rho\sigma}
  =
  \frac{1}{16\pi^2}
  \epsilon^{\mu\nu\rho\sigma}F_{\mu\nu}F_{\rho\sigma},
\]
which is the coefficient in \eqref{eq:abelian-axial-anomaly}; a field of
charge \(eq\) contributes the additional factor \(e^2q^2\).

\section{Measure Derivation}

The same coefficient is obtained from the Berezinian change of variables in a
regulated fermion path integral.  The construction uses the finite-dimensional
Berezin calculus of Chapter~\ref{ch:spinor-fields-fermionic-path-integrals};
the object that transforms is an ordered Berezinian density on a finite set of
odd coefficients, not a Borel measure on a set of fermion fields.

Work first on a compact Euclidean spin four-manifold, or in a finite box with
elliptic boundary conditions, so that the Euclidean Dirac operator
\[
  \mathcal D_A=\gamma^\mu(\partial_\mu-\ii A_\mu)
\]
has a discrete \(L^2\)-orthonormal eigenbasis after the infrared regulator has
been fixed.  The local formula obtained below is independent of this
infrared auxiliary choice.  Let \(P_\Lambda\) be the spectral projection onto
the span
\[
  V_\Lambda
  =
  \operatorname{span}\{\varphi_n:\ |\lambda_n|\le \Lambda\},
  \qquad
  \mathcal D_A\varphi_n=\lambda_n\varphi_n .
\]
The regulated fermion variables are odd coordinates
\[
  \psi_\Lambda(x)=\sum_{n\in I_\Lambda} c_n\varphi_n(x),
  \qquad
  \bar\psi_\Lambda(x)=\sum_{n\in I_\Lambda}\bar c_n\varphi_n^\dagger(x),
\]
with ordered Berezinian density
\[
  \Omega_\Lambda
  =
  \prod_{n\in I_\Lambda} d\bar c_n\,dc_n .
\]
An infinitesimal local chiral rotation
\[
  \psi(x)\mapsto \ee^{\ii\beta(x)\gamma_5}\psi(x),
  \qquad
  \bar\psi(x)\mapsto \bar\psi(x)\ee^{\ii\beta(x)\gamma_5},
\]
acts on \(V_\Lambda\) by the finite matrix
\[
  U_{\Lambda,\beta}
  =
  1+\ii P_\Lambda\beta\gamma_5P_\Lambda+O(\beta^2),
\]
where \(\beta\) denotes the multiplication operator by the smooth function
\(\beta(x)\).  The inverse determinant in the odd Berezinian transformation
law gives, with the convention that \(J_\Lambda\) is the factor multiplying
the transformed finite-dimensional Berezin integral,
\[
  \log J_\Lambda
  =
  2\ii\,
  \operatorname{tr}_{V_\Lambda}
  (P_\Lambda\beta\gamma_5P_\Lambda)
  +O(\beta^2).
\]
The factor of \(2\) records that \(\psi\) and \(\bar\psi\) are rotated with
the same chiral sign.

The sharp projection \(P_\Lambda\) may be replaced, for the local asymptotic
calculation, by the smooth gauge-covariant spectral regulator
\[
  \exp(-\mathcal D_A^2/M^2).
\]
This is the heat-kernel regulator of the Berezinian trace.  It introduces no
regulator field; \(M\) is the ultraviolet scale in the heat semigroup of the
elliptic operator \(\mathcal D_A^2\).  In the terminology of
Table~\ref{tab:regulator-integration-status-catalog}, this is a spectral
regulator for an operator trace, not a Pauli--Villars auxiliary-field or
determinant-ratio regulator.  The regulated logarithm used below is
\begin{equation}
  \log J_M
  =
  2\ii\lim_{M\to\infty}
  \operatorname{Tr}
  \left[
  \beta\,\gamma_5
  \exp\left(-\mathcal D_A^2/M^2\right)
  \right],
  \label{eq:fujikawa-regulated-trace}
\end{equation}
where \(\operatorname{Tr}\) is the trace over spacetime, spin, and internal
indices.  Equation \eqref{eq:fujikawa-regulated-trace} is therefore a
regulated finite-mode Berezinian statement followed by a heat-kernel
asymptotic limit.  It is the measure-language counterpart of the
dimensionally regulated triangle computation above, in the scheme where the
vector gauge Ward identity is preserved and the axial Ward identity carries
the local term.

The heat-kernel expansion gives the local limit
\begin{equation}
  \lim_{M\to\infty}
  \operatorname{tr}_{\mathrm{spin}}
  \left[
  \gamma_5
  \exp\left(-\mathcal D_A^2/M^2\right)
  \right](x,x)
  =
  \frac{1}{32\pi^2}
  \epsilon^{\mu\nu\rho\sigma}
  \operatorname{tr}_{R}(F_{\mu\nu}F_{\rho\sigma}),
  \label{eq:heat-kernel-anomaly-density}
\end{equation}
with \(\operatorname{tr}_{R}\) the trace in the fermion representation.  The
anomaly density is therefore the heat-kernel limit of the regulated
Berezinian traces of \(\gamma_5\) over cutoff eigenspaces.  Finite-dimensional
cyclic trace manipulations at fixed cutoff do not remove
\eqref{eq:heat-kernel-anomaly-density}, because the regulator is part of the
definition of the composite current.

The local heat-kernel density becomes a Fredholm index after the following
global data are specified.  The analytic input is the spin Dirac case of the
Atiyah--Singer index theorem.\footnote{M. F. Atiyah and
I. M. Singer, ``The index of elliptic operators. I,'' Ann. of Math. (2) 87
(1968), 484--530.  The boundary theorem used below is M. F. Atiyah,
V. K. Patodi, and I. M. Singer, ``Spectral asymmetry and Riemannian
geometry. I,'' Math. Proc. Cambridge Philos. Soc. 77 (1975), 43--69, with
Parts II and III in Math. Proc. Cambridge Philos. Soc. 78 (1975), 405--432
and 79 (1976), 71--99.}

\begin{quotedtheorem}[Spin Dirac index theorem, four-dimensional form]
\label{qthm:spin-dirac-index-four-dimensional}
Let \(M\) be a smooth compact oriented Riemannian spin four-manifold without
boundary.  Let \(S=S^+\oplus S^-\) be its complex spinor bundle, let
\(E_R\to M\) be a smooth Hermitian vector bundle associated to the fermion
representation \(R\), and let \(\nabla^E\) be a smooth unitary connection with
curvature \(F_R\) in the same gauge convention as the local anomaly
calculation.  The chiral Dirac operator
\[
  \mathcal D_{A,+}:
  \Gamma(S^+\otimes E_R)\longrightarrow
  \Gamma(S^-\otimes E_R)
\]
is elliptic and hence has finite-dimensional kernel and cokernel.  Its
Fredholm index is
\begin{equation}
  \operatorname{index}\mathcal D_{A,+}
  =
  \dim\ker\mathcal D_{A,+}
  -
  \dim\ker\mathcal D_{A,-}
  =
  \int_M
  \bigl[\widehat A(TM)\operatorname{ch}(E_R,\nabla^E)\bigr]_4 ,
  \label{eq:dirac-index-as-form}
\end{equation}
where \([\cdots]_4\) denotes the four-form part of the Chern--Weil form,
\(\widehat A(TM)\) is the \(\widehat A\)-class of the tangent bundle, and
\(\operatorname{ch}(E_R,\nabla^E)\) is the Chern character form of the gauge
bundle.  The operator \(\mathcal D_{A,-}\) is the opposite-chirality Dirac
operator
\(\Gamma(S^-\otimes E_R)\to\Gamma(S^+\otimes E_R)\).
\end{quotedtheorem}

\paragraph{Index mechanism and four-form normalization.}
The quoted theorem imports the global equality between a Fredholm index and a
characteristic number.  The part used by the anomaly calculation can be made
more explicit.  Let
\[
  \mathcal D_A
  =
  \begin{pmatrix}
    0&\mathcal D_{A,-}\\
    \mathcal D_{A,+}&0
  \end{pmatrix}
\]
be the self-adjoint Dirac operator on \(S\otimes E_R\), and let
\(\Gamma_5\) denote the Euclidean chirality operator, acting by \(+1\) on
\(S^+\) and by \(-1\) on \(S^-\).  For every \(t>0\), elliptic heat-kernel
theory makes \(\Gamma_5\ee^{-t\mathcal D_A^2}\) trace class, and the
McKean--Singer cancellation gives
\begin{equation}
  \operatorname{Tr}\!\left(\Gamma_5\ee^{-t\mathcal D_A^2}\right)
  =
  \dim\ker\mathcal D_{A,+}
  -
  \dim\ker\mathcal D_{A,-}.
  \label{eq:dirac-index-mckean-singer-anomaly}
\end{equation}
Indeed, every positive eigenvector of
\(\mathcal D_{A,-}\mathcal D_{A,+}\) is paired by \(\mathcal D_{A,+}\) with a
positive eigenvector of \(\mathcal D_{A,+}\mathcal D_{A,-}\), with the same
eigenvalue and opposite chirality contribution.  The heat supertrace is
therefore independent of \(t\), and the index is the small-\(t\) integral of
the local supertrace coefficient.

The local coefficient is the analytic part of the index theorem.  In the
heat-kernel proof it is obtained from the Lichnerowicz formula, Chern--Weil
invariance, and Getzler rescaling of the Clifford variables.  The pure gauge
coefficient can be checked locally, and this check fixes the convention used
in the anomaly density.  Choose normal coordinates at a point \(x\), trivialize
\(E_R\) so that the connection one-form vanishes at \(x\), and write
\(\Gamma^{\mu\nu}=\frac12[\Gamma^\mu,\Gamma^\nu]\).  The gauge-curvature
endomorphism in the square of the Dirac operator is
\[
  \mathcal E_F(x)
  =
  \frac12\Gamma^{\mu\nu}F_{R,\mu\nu}(x).
\]
Terms with fewer than four Clifford matrices have zero
\(\Gamma_5\)-supertrace.  The first pure gauge term that survives in the
diagonal heat-kernel parametrix is therefore
\[
  (4\pi t)^{-2}
  \frac{t^2}{2}
  \operatorname{tr}_{\mathrm{spin},R}
  \left(\Gamma_5\mathcal E_F^2\right).
\]
Using
\[
  \operatorname{tr}_{\mathrm{spin}}
  \left(\Gamma_5\Gamma^{\mu\nu}\Gamma^{\rho\sigma}\right)
  =
  4\epsilon^{\mu\nu\rho\sigma},
  \qquad \epsilon^{0123}=+1,
\]
one obtains
\[
  (4\pi)^{-2}\frac12\frac14\,
  4\epsilon^{\mu\nu\rho\sigma}
  \operatorname{tr}_R(F_{R,\mu\nu}F_{R,\rho\sigma})
  =
  \frac{1}{32\pi^2}
  \epsilon^{\mu\nu\rho\sigma}
  \operatorname{tr}_R(F_{R,\mu\nu}F_{R,\rho\sigma}).
\]
The spin-curvature terms in the same Getzler expansion assemble with the
pure gauge term into the Chern--Weil form
\(\widehat A(TM)\operatorname{ch}(E_R,\nabla^E)\).  With the chapter's
anti-Hermitian curvature convention
\(\mathsf F_R=-\ii F_R\), the Chern character appearing in
Theorem~\ref{qthm:spin-dirac-index-four-dimensional} is
\[
  \operatorname{ch}(E_R,\nabla^E)
  =
  \operatorname{tr}_R
  \exp\!\left(\frac{\ii\mathsf F_R}{2\pi}\right)
  =
  \operatorname{tr}_R
  \exp\!\left(\frac{F_R}{2\pi}\right).
\]
Consequently the gauge part of the four-form is
\begin{equation}
  \bigl[\operatorname{ch}(E_R,\nabla^E)\bigr]_4
  =
  \frac1{2(2\pi)^2}\operatorname{tr}_R(F_R\wedge F_R)
  =
  \frac1{32\pi^2}
  \epsilon^{\mu\nu\rho\sigma}
  \operatorname{tr}_R(F_{\mu\nu}F_{\rho\sigma})\,\dd^4x .
  \label{eq:dirac-index-four-form-normalization}
\end{equation}
The first equality is Chern--Weil algebra; the second uses
\(F_R=\frac12F_{\mu\nu}\dd x^\mu\wedge\dd x^\nu\) and the orientation
\(\epsilon^{0123}=+1\).  This explains why the local Berezinian density
\eqref{eq:heat-kernel-anomaly-density} and the global Fredholm index have the
same coefficient once the closed spin hypotheses of the theorem are imposed.

On a flat four-torus or on \(\mathbb R^4\) after compactification by a smooth
finite-action connection for which no boundary contribution is present, the
tangent-bundle factor contributes only \(1\), and the gauge part of
\eqref{eq:dirac-index-as-form} becomes the density computed in
\eqref{eq:heat-kernel-anomaly-density}:
\begin{equation}
  \operatorname{index}\mathcal D_{A,+}
  =
  n_+-n_-
  =
  \frac1{32\pi^2}
  \int_M
  \epsilon^{\mu\nu\rho\sigma}
  \operatorname{tr}_{R}(F_{\mu\nu}F_{\rho\sigma})\,\dd^4x .
  \label{eq:dirac-index-anomaly}
\end{equation}
Here \(n_+=\dim\ker\mathcal D_{A,+}\) and
\(n_-=\dim\ker\mathcal D_{A,-}\), equivalently the numbers of zero modes of
chirality \(+1\) and \(-1\) in the convention of this chapter.  The equality
between the integrated heat-kernel density and this integer uses the global
hypotheses in the theorem.  Without those hypotheses the local Ward identity
still has the density \eqref{eq:heat-kernel-anomaly-density}, but its integral
is not by itself a Fredholm index.

If \(M\) has a boundary, assume first that the Riemannian metric and the gauge
connection have product form on a collar \(u\in[0,\varepsilon)\) of
\(\partial M\).  In this collar the chiral Dirac operator can be written
\[
  \mathcal D_A
  =
  \gamma(n)\bigl(\partial_u+B_A\bigr),
\]
where \(n\) is the outward unit normal and \(B_A\) is the self-adjoint
tangential Dirac operator on \(\partial M\).  Let
\[
  \eta_{B_A}(s)
  =
  \sum_{\lambda\in\operatorname{Spec}(B_A)\setminus\{0\}}
  \operatorname{sign}(\lambda)|\lambda|^{-s}
\]
for \(\operatorname{Re}s\) large, continued meromorphically to \(s=0\), and
let \(h_{B_A}=\dim\ker B_A\).  With the APS boundary condition
\(P_{\geq0}\psi|_{\partial M}=0\), where \(P_{\geq0}\) is the spectral
projection of \(B_A\) onto nonnegative eigenvalues, the Atiyah--Patodi--Singer
theorem gives
\begin{equation}
  \operatorname{index}\mathcal D^{\mathrm{APS}}_{A,+}
  =
  \int_M
  \bigl[\widehat A(TM)\operatorname{ch}(E_R,\nabla^E)\bigr]_4
  -
  \frac{\eta_{B_A}(0)+h_{B_A}}{2}.
  \label{eq:aps-index-boundary}
\end{equation}
For non-product data one either deforms to product form near the boundary, or
includes the corresponding local transgression term; the sum of the bulk,
local boundary, and spectral terms is the invariant integer.  The appearance
of \(\eta_{B_A}\) shows that boundary or interface anomaly data include
spectral terms in addition to the bulk density.  Boundary extension,
global-anomaly, and inflow applications use \eqref{eq:aps-index-boundary};
the local Ward identity in the present chapter uses only the bulk density
before this spectral correction is imposed.

\section{Nonabelian Flavor and Gauge Anomalies}

\begin{definition}[Trace anomaly coefficient]
\label{def:trace-anomaly-coefficient}
For several Weyl fermions, let \(T^A\) be a generator of a global symmetry
current and let \(t^a,t^b\) be generators of gauge or background gauge fields.
The Hermitian generator and curvature conventions are those stated in
\eqref{eq:ym-lie-algebra-conventions}, \eqref{eq:ym-trace-kappa}, and
\eqref{eq:ym-lagrangian-density}; in particular the default basis has
\(\operatorname{tr}(t^a t^b)=\delta^{ab}\).  The mixed anomaly coefficient is
\begin{equation}
  \mathcal A^{Aab}
  =
  \frac12
  \left[
  \operatorname{tr}_{\mathrm{left}}
  \bigl(T^A\{t^a,t^b\}\bigr)
  -
  \operatorname{tr}_{\mathrm{right}}
  \bigl(T^A\{t^a,t^b\}\bigr)
  \right].
  \label{eq:flavor-anomaly-coefficient}
\end{equation}
With background field strengths \(F^a_{\mu\nu}\), the corresponding current
identity has the form
\begin{equation}
  \partial_\mu J_A^\mu
  =
  \frac1{32\pi^2}
  \mathcal A^{Aab}
  \epsilon^{\mu\nu\rho\sigma}
  F^a_{\mu\nu}F^b_{\rho\sigma}
  +
  \text{explicit breaking terms}.
  \label{eq:nonabelian-flavor-anomaly}
\end{equation}
The factor \(1/2\) in \eqref{eq:flavor-anomaly-coefficient} is part of the
definition of \(\mathcal A^{Aab}\); it compensates for the two identical vector
insertions in the anticommutator.  For a single Dirac fermion with an axial
generator acting by \(+1\) on the left-handed component and \(-1\) on the
right-handed component, and with a unit Abelian vector charge, this definition
gives \(\mathcal A=2\), so \eqref{eq:nonabelian-flavor-anomaly} reduces to the
coefficient \(1/(16\pi^2)\) in \eqref{eq:abelian-axial-anomaly}.
\end{definition}

\begin{proposition}[Cubic gauge obstruction]
\label{prop:cubic-gauge-anomaly-obstruction}
For left-handed Weyl fermions in representations \(R_i\) of a gauge group,
the cubic part of the one-loop local consistent gauge anomaly vanishes if and
only if
\begin{equation}
  \sum_i
  \operatorname{tr}_{R_i}
  \bigl(t^a\{t^b,t^c\}\bigr)
  =
  0
  \qquad
  \text{for all }a,b,c.
  \label{eq:gauge-anomaly-cancellation}
\end{equation}
If this tensor is nonzero for a dynamical gauge symmetry, no local
counterterm can make the corresponding one-loop effective action satisfy the
Slavnov--Taylor identity in that background-field chart.  If the tensor
vanishes, this cubic local obstruction is absent; maintaining the full
Slavnov--Taylor identities then reduces to the absence of the remaining
BRST-anomaly classes in the chosen theory and regulator scheme.
\end{proposition}

\begin{proof}[Proof from local descent]
The obstruction is computed in the perturbative local BRST bicomplex on a
contractible coordinate chart of the bundle of background connections.  The
anti-Hermitian curvature \(\mathsf F=\mathsf F^a t^a\) is an even two-form,
so the exterior products \(\mathsf F^a\mathsf F^b\mathsf F^c\) are symmetric
under interchange of \(a,b,c\).  Therefore the degree-three invariant
polynomial of one left-handed Weyl fermion depends on the representation only
through the symmetric cubic tensor:
\[
  \operatorname{tr}_{R}(\mathsf F^3)
  =
  \operatorname{tr}_{R}(t^{(a}t^bt^{c)})
  \mathsf F^a\mathsf F^b\mathsf F^c .
\]
Equivalently, after one gauge variation has been converted into a ghost and
two curvatures are read as external gauge-field insertions, the same tensor is
written as
\[
  \operatorname{tr}_{R}\bigl(t^a\{t^b,t^c\}\bigr)
\]
up to the conventional normalization already fixed in
\eqref{eq:flavor-anomaly-coefficient}.  Right-handed Weyl fermions contribute
with the opposite sign because the chirality projector entering the regulated
fermion determinant is reversed.

The descent map sends the invariant polynomial to a local representative
\(I_4^{(1)}\in\Omega_{\rm loc}^{4,1}\):
\[
  \operatorname{tr}_{R}(\mathsf F^3)=\dd I_5^{(0)},
  \qquad
  s I_5^{(0)}=\dd I_4^{(1)} .
\]
Changing the subtraction convention for the one-loop effective action adds a
local counterterm \(B_4^{(0)}\), and the anomaly representative shifts by
\[
  I_4^{(1)}
  \longmapsto
  I_4^{(1)}+sB_4^{(0)}+\dd B_3^{(1)} .
\]
Thus counterterms change representatives inside the same class of
\(H^{1,4}(s\mid\dd)\).  The cohomological input used here is the local
BRST-classification theorem: for four-dimensional chiral Yang--Mills
theories in this polynomial local setting, the semisimple cubic gauge part of
\(H^{1,4}(s\mid\dd)\) is generated by the descent classes of the degree-three
invariant polynomials \(\operatorname{tr}_{R}(\mathsf F^3)\).  The vanishing
of the total symmetric cubic tensor therefore removes precisely this local
class, while a nonzero tensor gives a class outside the image of local
counterterm variations.

The non-removability of a nonzero cubic tensor can also be seen without using
the full local BRST classification.  Restrict the background to a maximal
torus \(T\) with Lie algebra \(\mathfrak t\).  Since an invariant polynomial
on a complex semisimple Lie algebra is determined by its restriction to a
Cartan subalgebra, a nonzero cubic invariant has a nonzero restriction to
\(\mathfrak t\).  Indeed, regular semisimple elements are conjugate to
\(\mathfrak t\) and form a Zariski-dense subset; an invariant polynomial that
vanishes on \(\mathfrak t\) therefore vanishes on all regular semisimple
elements and hence identically.  Thus a nonzero semisimple cubic gauge
anomaly can be detected after Abelianizing to commuting gauge fields
\[
  A=A^i H_i,\qquad F^i=\dd A^i .
\]
On this Abelianized chart the cubic anomaly representative has the form
\[
  \int_M C_{i;jk}\lambda^i F^jF^k,
  \qquad C_{i;jk}=C_{i;kj},
\]
and its invariant cubic coordinate is the completely symmetric part of
\(C_{i;jk}\).  The most general cubic local counterterm that can change this
\(\lambda F F\) coefficient has, modulo gauge-invariant terms and total
derivatives, the form
\[
  B_4
  =
  \frac12 H_{ij,k}A^iA^jF^k,
  \qquad H_{ij,k}=-H_{ji,k}.
\]
Using \(\delta_\lambda A^i=\dd\lambda^i\), \(\delta_\lambda F^i=0\), and
\(\dd F^i=0\), its variation on a closed spacetime is
\[
  \delta_\lambda\!\int_M B_4
  =
  -\int_M
  H_{ij,k}\lambda^i F^jF^k .
\]
Since \(F^jF^k=F^kF^j\), the effective shift of the coefficient symmetric in
the two curvature labels is
\[
  \Delta C_{i;jk}
  =
  -\frac12\bigl(H_{ij,k}+H_{ik,j}\bigr).
\]
The complete symmetrization over \(i,j,k\) vanishes identically because
\(H\) is antisymmetric in its first two labels.  Therefore local
counterterms cannot change the fully symmetric Cartan coefficient of the
Abelianized anomaly.  By the Cartan-restriction argument above, a nonzero
semisimple cubic invariant tensor cannot be removed by a local counterterm.
This elementary Abelianized calculation is the part of the obstruction that
the anomaly-matching and WZW sections use repeatedly.

The cubic gauge part of the anomaly polynomial for a left-handed Weyl fermion
in \(R_i\) is proportional to
\(\operatorname{tr}_{R_i}(\mathsf F^3)\).  Expanding
\(\mathsf F=\mathsf F^a t^a\), the coefficient of
\(\mathsf F^a\mathsf F^b\mathsf F^c\) is the completely symmetric tensor
\(\operatorname{tr}_{R_i}(t^{(a}t^b t^{c)})\), equivalently
\(\frac16\operatorname{tr}_{R_i}\) summed over all permutations.  In component
Ward identities the same tensor is written as
\(\operatorname{tr}_{R_i}(t^a\{t^b,t^c\})\), with the conventional factor
absorbed into the anomaly coefficient.  Summing over fermions gives
\eqref{eq:gauge-anomaly-cancellation}.

A local counterterm changes the anomaly representative by a BRST coboundary.
The Abelianized counterterm calculation above already proves that a nonzero
semisimple cubic invariant tensor cannot be removed by changing local
counterterms.  The remaining use of local BRST cohomology is the converse
classification statement: in the polynomial local perturbative Yang--Mills
setting, once this invariant-polynomial coefficient vanishes, the semisimple
cubic gauge-anomaly class itself is absent.  Other anomaly classes, global
anomaly holonomies, and regulator-specific Slavnov--Taylor restoration
conditions are separate data.  The last sentence is therefore not an
all-purpose nonperturbative existence theorem; it is the perturbative BRST
consistency reduction in a specified gauge-fixed regularized framework.
\end{proof}

\section{Consistency Condition and Descent}

The anomaly of a gauge symmetry is constrained before any coefficient is
computed.  Let \(W[A]\) be the renormalized effective action for a background
gauge field \(A\).  Under an infinitesimal gauge transformation with parameter
\(\zeta=\zeta^a t^a\), write
\[
  \delta_\zeta W[A]=\mathcal A(\zeta,A).
\]
Here the infinitesimal action on a functional \(F[A]\) is the derivation
\begin{equation}
  \delta_\zeta F[A]
  =
  \int_M \dd^4x\,
  \frac{\delta F}{\delta A_\mu^a(x)}
  \,(D_\mu\zeta)^a(x),
  \qquad
  D_\mu\zeta=\partial_\mu\zeta+[A_\mu,\zeta]_{\mathrm H},
  \label{eq:gauge-variation-functional-derivation}
\end{equation}
understood either after introducing a regulator that makes the functional
derivatives ordinary finite-dimensional derivatives, or in the formal local
functional calculus.  The gauge parameters are external test functions and
are not varied by \(\delta_{\zeta_i}\).  With these conventions the
variations close on the connection:
\begin{equation}
  [\delta_{\zeta_1},\delta_{\zeta_2}]A_\mu
  =
  D_\mu[\zeta_1,\zeta_2]_{\mathrm H}.
  \label{eq:gauge-variation-closure-on-connection}
\end{equation}
Indeed,
\[
  \delta_{\zeta_1}\delta_{\zeta_2}A_\mu
  -
  \delta_{\zeta_2}\delta_{\zeta_1}A_\mu
  =
  [D_\mu\zeta_1,\zeta_2]_{\mathrm H}
  -
  [D_\mu\zeta_2,\zeta_1]_{\mathrm H},
\]
and expanding \(D_\mu\) gives
\[
  [\partial_\mu\zeta_1,\zeta_2]_{\mathrm H}
  +
  [\zeta_1,\partial_\mu\zeta_2]_{\mathrm H}
  +
  [[A_\mu,\zeta_1]_{\mathrm H},\zeta_2]_{\mathrm H}
  +
  [\zeta_1,[A_\mu,\zeta_2]_{\mathrm H}]_{\mathrm H}
  =
  \partial_\mu[\zeta_1,\zeta_2]_{\mathrm H}
  +[A_\mu,[\zeta_1,\zeta_2]_{\mathrm H}]_{\mathrm H},
\]
where the last equality is the Jacobi identity.  Equation
\eqref{eq:gauge-variation-closure-on-connection} implies, by the chain rule,
\[
  [\delta_{\zeta_1},\delta_{\zeta_2}]F[A]
  =
  \delta_{[\zeta_1,\zeta_2]_{\mathrm H}}F[A]
\]
for every functional \(F\) in the same differentiability class.
Applying this identity to \(W[A]\), and using
\(\mathcal A(\zeta,A)=\delta_\zeta W[A]\), gives
\begin{equation}
  \delta_{\zeta_1}\mathcal A(\zeta_2,A)
  -
  \delta_{\zeta_2}\mathcal A(\zeta_1,A)
  =
  \mathcal A([\zeta_1,\zeta_2]_{\mathrm H},A).
  \label{eq:wess-zumino-consistency}
\end{equation}
This is the Wess--Zumino consistency condition.  It is an integrability
condition: a proposed local expression for \(\mathcal A\) is the gauge
variation of a single effective action only if its first variations satisfy
the gauge-algebra commutator above.  In BRST language, replace \(\zeta\) by
the ghost \(c\).  The anomaly is a local ghost-number-one functional
\(\mathcal A(c,A)\) obeying
\begin{equation}
  s\mathcal A=0,
  \label{eq:brst-anomaly-cocycle}
\end{equation}
and adding a local counterterm \(X[A]\) shifts it by
\[
  \mathcal A\longmapsto \mathcal A+sX .
\]
Thus the possible local gauge anomalies are cohomology classes of the BRST
differential at ghost number one.

The local form of this statement is the relative cohomology described in
Chapter~\ref{ch:volume-ii-18}.  Let
\[
  \Omega_{\mathrm{loc}}^{p,q}
\]
be the vector space of local \(p\)-forms of ghost number \(q\), built from the
background connection, the ghost, and finitely many derivatives, with
coefficients that are polynomial or formal power series in the fields in the
chosen local coordinate chart.  In the abstract cohomological notation
\(H(s\mid d)\), write \(d\) for the spacetime exterior derivative; in the
connection formulas below the same exterior derivative is denoted \(\dd\).
The derivative \(d\) has bidegree \((1,0)\),
\[
  d:\Omega_{\mathrm{loc}}^{p,q}\longrightarrow
  \Omega_{\mathrm{loc}}^{p+1,q},
\]
and the BRST differential has bidegree \((0,1)\),
\[
  s:\Omega_{\mathrm{loc}}^{p,q}\longrightarrow
  \Omega_{\mathrm{loc}}^{p,q+1}.
\]
Both are odd with respect to total degree, and with the sign convention used
here
\begin{equation}
  d^2=0,\qquad s^2=0,\qquad sd+ds=0 .
  \label{eq:descent-bicomplex-differentials}
\end{equation}
Thus \(\Omega_{\mathrm{loc}}^{\bullet,\bullet}\) is a bicomplex.  For a
four-dimensional integrated anomaly, choose a representative
\[
  a_4^{(1)}\in\Omega_{\mathrm{loc}}^{4,1}.
\]
It is a relative cocycle when there exists
\(a_3^{(2)}\in\Omega_{\mathrm{loc}}^{3,2}\) such that
\begin{equation}
  s a_4^{(1)}-d a_3^{(2)}=0 .
  \label{eq:relative-anomaly-cocycle}
\end{equation}
Two representatives define the same relative class when there exist
\(b_4^{(0)}\in\Omega_{\mathrm{loc}}^{4,0}\) and
\(b_3^{(1)}\in\Omega_{\mathrm{loc}}^{3,1}\) such that
\begin{equation}
  a_4^{(1)}
  \sim
  a_4^{(1)}+s b_4^{(0)}+d b_3^{(1)} .
  \label{eq:relative-anomaly-coboundary}
\end{equation}
The quotient of cocycles by this equivalence relation is
\[
  H^{1,4}(s\mid d).
\]

\begin{proposition}[Local anomalies as relative BRST classes]
\label{prop:local-anomalies-relative-brst-classes}
At the level of local functionals, a four-dimensional consistent gauge anomaly
with local density \(a_4^{(1)}\) determines a class
\([a_4^{(1)}]\in H^{1,4}(s\mid d)\).  Changing the renormalized effective
action by a local counterterm changes \(a_4^{(1)}\) by the coboundary
\eqref{eq:relative-anomaly-coboundary}.  Conversely, a representative of
\(H^{1,4}(s\mid d)\) defines a local functional whose BRST variation obeys
the infinitesimal Wess--Zumino consistency condition.
\end{proposition}

\begin{proof}
Write the anomalous variation as
\[
  \delta_\zeta W[A]=\int_M a_4^{(1)}(\zeta,A).
\]
Replacing the even parameter \(\zeta\) by the odd ghost gives
\(\int_M a_4^{(1)}(c,A)\).  The BRST differential \(s\) is the
Chevalley--Eilenberg differential for infinitesimal gauge transformations,
with the field-dependent term in \(s c=-\frac12[c,c]\) included.  Equation
\eqref{eq:wess-zumino-consistency} is precisely the statement that this
ghost-number-one integrated functional is a Chevalley--Eilenberg cocycle,
hence
\[
  s\int_M a_4^{(1)}=0 .
\]
For a local density this means that \(s a_4^{(1)}\) integrates to zero for all
compactly supported field and ghost configurations.  In the local variational
bicomplex, the algebraic Poincare lemma identifies such local top forms with
total derivatives modulo globally topological terms fixed separately by the
chosen bundle sector.  Thus there exists
\(a_3^{(2)}\in\Omega_{\rm loc}^{3,2}\) with
\[
  s a_4^{(1)}=d a_3^{(2)},
\]
which is \eqref{eq:relative-anomaly-cocycle}.  If \(W[A]\) is changed by a
local counterterm \(\int_M b_4^{(0)}\), its anomalous variation changes by
\[
  s\int_M b_4^{(0)}
  =
  \int_M s b_4^{(0)} ,
\]
and adding an identically integrated total derivative changes the local
representative by \(d b_3^{(1)}\).  This is exactly the equivalence
\eqref{eq:relative-anomaly-coboundary}; it is the statement that a finite
local counterterm changes the representative but not the relative cohomology
class.

Conversely, if \eqref{eq:relative-anomaly-cocycle} holds, then on a closed
spacetime, or with boundary conditions for which the boundary integral is part
of the specified inflow problem, the integrated BRST variation of
\(\int_M a_4^{(1)}\) vanishes.  Expanding \(s\) in ghosts recovers the
antisymmetrized commutator of two ordinary gauge variations, while
\(s c=-\frac12[c,c]\) supplies the term with gauge parameter
\([\zeta_1,\zeta_2]\).  This is exactly the infinitesimal
Wess--Zumino consistency condition \eqref{eq:wess-zumino-consistency}.
\end{proof}

\paragraph{Finite-regulator scheme meaning of the class.}
The relative class above is the infinitesimal local shadow of a finite
regulator statement.  Let \(\mathcal B_\Lambda\) be a regulated space of
background fields, let \(\mathcal G_\Lambda\) be the regulated background
gauge group acting on the right, \(A\mapsto A^g\), and suppose a local chart
of the regulated determinant line has been chosen so that the fermion
functional is written as a scalar \(W_\Lambda[A]\).  Its finite anomaly
cochain is
\begin{equation}
  C_\Lambda(g;A)
  =
  W_\Lambda[A^g]-W_\Lambda[A].
  \label{eq:finite-regulator-anomaly-cochain}
\end{equation}
At this stage the statement is an identity in the regulated chart; continuum
content enters after locality, convergence, and counterterm estimates have
been supplied.
Functoriality of the gauge action gives the exact groupoid cocycle equation
\begin{equation}
  C_\Lambda(gh;A)
  =
  C_\Lambda(h;A^g)+C_\Lambda(g;A),
  \label{eq:finite-regulator-anomaly-cocycle}
\end{equation}
with the order fixed by \((A^g)^h=A^{gh}\).  Differentiating
\eqref{eq:finite-regulator-anomaly-cocycle} at
\(g=\exp(t\zeta_1)\), \(h=\exp(t\zeta_2)\), and then taking the local
continuum representative gives the Wess--Zumino condition
\eqref{eq:wess-zumino-consistency}; replacing the even parameters by the
ghost gives the relative BRST cocycle
\eqref{eq:relative-anomaly-cocycle}.

Changing the regulated Lagrangian by a finite local counterterm
\(B_\Lambda[A]\), or choosing a determinant-line frame whose scalar logarithm
differs by \(B_\Lambda[A]\), replaces \(W_\Lambda\) by
\(W_\Lambda+B_\Lambda\) and hence
\begin{equation}
  C_\Lambda(g;A)
  \longmapsto
  C_\Lambda(g;A)+B_\Lambda[A^g]-B_\Lambda[A].
  \label{eq:finite-regulator-anomaly-coboundary}
\end{equation}
The infinitesimal local density changes by
\(a_4^{(1)}\mapsto a_4^{(1)}+s b_4^{(0)}+d b_3^{(1)}\), exactly as in
\eqref{eq:relative-anomaly-coboundary}.  Thus regulator and subtraction
choices that remain within the same local background-field scheme change the
representative, not the relative cohomology class.  A nonzero local class
means that no local counterterm in that perturbative gauge-fixed scheme can
turn the regulated determinant chart into a gauge-invariant scalar effective
action.  The statement is local and perturbative: vanishing of this class is
the condition that removes this obstruction, while an order-by-order
construction of a gauge theory must still build the counterterms and verify
the Slavnov--Taylor or BV quantum master identities in the chosen regulator.
Flat global holonomies of the determinant or Pfaffian line are separate
finite-background data and are treated in
Chapter~\ref{ch:curved-eta-invariants-global-anomalies}.

For four-dimensional chiral fermions, this relative cohomology class is
represented by the descent of a six-form anomaly polynomial.  To state the
normalization without hiding convention-dependent factors, introduce the
anti-Hermitian connection
\[
  \mathsf A=-\ii A,
  \qquad
  \mathsf F=\dd\mathsf A+\mathsf A^2=-\ii F ,
\]
where \(A\) and \(F\) are the Hermitian gauge potential and curvature used in
the Yang--Mills chapters.  Let \(\mathsf c\) be the corresponding odd
anti-Hermitian ghost.  On the descent variables,
\begin{equation}
  s\mathsf A=\dd\mathsf c+[\mathsf A,\mathsf c],
  \qquad
  s\mathsf c=-\mathsf c^2,
  \qquad
  s\mathsf F=[\mathsf F,\mathsf c].
  \label{eq:descent-brst-connection}
\end{equation}
Products of differential forms below include both matrix multiplication and
wedge product with total-degree Koszul signs.  For a left-handed Weyl fermion
in a representation \(R\), define the six-form
\[
  I_6^R
  =
  \kappa\,\operatorname{tr}_R(\mathsf F^3),
  \qquad
  \kappa=\frac{\ii}{24\pi^2}
\]
in the Euclidean effective-action convention.  Removing the overall Euclidean
phase gives the corresponding Lorentzian Ward-identity normalization.  A
right-handed Weyl fermion contributes with the opposite sign.  For a set of
left- and right-handed Weyl fermions,
\[
  I_6
  =
  \kappa
  \left[
    \sum_{i\in L}\operatorname{tr}_{R_i}(\mathsf F^3)
    -
    \sum_{j\in R}\operatorname{tr}_{R_j}(\mathsf F^3)
  \right].
\]

The Chern--Simons five-form \(I_5^{(0)}\) is any local form satisfying
\[
  I_6=\dd I_5^{(0)}.
\]
For one left-handed representation \(R\), a standard representative is
\[
  I_{5,R}^{(0)}
  =
  \kappa\,\operatorname{tr}_R
  \left(
    \mathsf A(\dd\mathsf A)^2
    +
    \frac32\mathsf A^3\dd\mathsf A
    +
    \frac35\mathsf A^5
  \right),
  \qquad
  \dd I_{5,R}^{(0)}=I_6^R .
\]
The numerical coefficients in this representative are fixed by the
Chern--Weil homotopy, not by convention.  Set
\[
  \mathsf A_t=t\mathsf A,\qquad
  \mathsf F_t=\dd\mathsf A_t+\mathsf A_t^2
  =t\,\dd\mathsf A+t^2\mathsf A^2 .
\]
For the invariant polynomial
\(P(X)=\kappa\,\operatorname{tr}_R(X^3)\), the Bianchi identity for
\(\mathsf F_t\) and cyclicity of the trace give
\[
  \frac{\dd}{\dd t}P(\mathsf F_t)
  =
  3\kappa\,\dd\,\operatorname{tr}_R(\mathsf A\,\mathsf F_t^2).
\]
Since \(\mathsf F_0=0\), integration in \(t\) gives
\[
  I_{5,R}^{(0)}
  =
  3\kappa\int_0^1
  \operatorname{tr}_R\!\left[
    \mathsf A
    \bigl(t\,\dd\mathsf A+t^2\mathsf A^2\bigr)^2
  \right]\dd t .
\]
Expanding the square and using cyclicity with the form degrees
\(\deg\mathsf A=1\), \(\deg\dd\mathsf A=2\), one obtains
\[
\begin{aligned}
  I_{5,R}^{(0)}
  =
  \kappa\,\operatorname{tr}_R
  \biggl[
    &3\!\int_0^1 t^2\dd t\,
      \mathsf A(\dd\mathsf A)^2
\\
    &+3\!\int_0^1 t^3\dd t\,
      \mathsf A\bigl((\dd\mathsf A)\mathsf A^2
      +\mathsf A^2\dd\mathsf A\bigr)
    +3\!\int_0^1 t^4\dd t\,\mathsf A^5
  \biggr]
\\
  =
  \kappa\,\operatorname{tr}_R
  \left(
    \mathsf A(\dd\mathsf A)^2
    +\frac32\mathsf A^3\dd\mathsf A
    +\frac35\mathsf A^5
  \right).
\end{aligned}
\]
Thus the factors \(1,\frac32,\frac35\) are the explicit transgression
integrals \(3\int_0^1t^2\dd t\),
\(3\int_0^1 2t^3\dd t\), and
\(3\int_0^1t^4\dd t\).  They use only the fixed trace convention and the
matrix-valued differential-form product.  A different path of connections or
a different Chern--Simons representative changes \(I_5^{(0)}\) by an exact
form and hence changes \(I_4^{(1)}\) only by a relative coboundary of the
kind already displayed in \eqref{eq:relative-anomaly-coboundary}.
Because \(sI_6=0\) and \(I_6=\dd I_5^{(0)}\),
\[
  \dd(sI_5^{(0)})=-s(\dd I_5^{(0)})=-sI_6=0 .
\]
The algebraic Poincare lemma on local forms then gives the first descent
equation
\begin{equation}
  s I_5^{(0)}
  =
  \dd I_4^{(1)}(\mathsf c,\mathsf A).
  \label{eq:first-brst-descent}
\end{equation}
Applying \(s\) again and using \eqref{eq:descent-bicomplex-differentials}
gives a second descent equation
\begin{equation}
  s I_4^{(1)}
  =
  \dd I_3^{(2)}
  \label{eq:second-brst-descent}
\end{equation}
for some \(I_3^{(2)}\in\Omega_{\mathrm{loc}}^{3,2}\).  Hence
\(I_4^{(1)}\) is a representative of \(H^{1,4}(s\mid d)\) in the convention
of \eqref{eq:relative-anomaly-cocycle}.  Replacing the ghost \(\mathsf c\) by
an even infinitesimal parameter \(\lambda\) gives the ordinary gauge variation
\[
  \delta_\lambda \mathsf A=\dd\lambda+[\mathsf A,\lambda],
  \qquad
  \delta_\lambda I_5^{(0)}
  =
  \dd I_4^{(1)}(\lambda,\mathsf A).
\]
In the same left-handed convention,
\[
  I_{4,R}^{(1)}(\mathsf c,\mathsf A)
  =
  \kappa\,\operatorname{tr}_R
  \left[
    \mathsf c\,\dd
    \left(
      \mathsf A\,\dd\mathsf A+\frac12\mathsf A^3
    \right)
  \right],
\]
up to adding \(\dd\)-exact terms and BRST-exact local counterterm variations.
The relative coefficient \(\frac12\) in this four-form is the corresponding
homotopy coefficient \(6\int_0^1(1-t)t^2\dd t=\frac12\); the coefficient of
\(\mathsf A\dd\mathsf A\) is \(6\int_0^1(1-t)t\dd t=1\).  This is the same
descent calculation as above, now evaluated on the vertical gauge direction
\(\delta_\lambda\mathsf A=\dd\lambda+[\mathsf A,\lambda]\).  The displayed
\(\mathsf c\,\dd(\cdots)\) representative is chosen because its integrated
form pairs directly with the effective-action variation
\eqref{eq:consistent-descent-anomaly-representative}; the equivalent
\(\dd\mathsf c\)-representatives differ from it by exact local forms.
The consistent anomaly is
\begin{equation}
  \delta_\lambda W[\mathsf A]
  =
  \int_M I_4^{(1)}(\lambda,\mathsf A).
  \label{eq:consistent-descent-anomaly-representative}
\end{equation}
The cancellation condition \eqref{eq:gauge-anomaly-cancellation} is precisely
the vanishing of the six-form polynomial \(I_6\) for the gauged symmetry, or
equivalently the vanishing of the symmetric cubic tensor
\[
  d^{abc}
  =
  \sum_{i\in L}\operatorname{tr}_{R_i}
  \bigl(t^{(a}t^b t^{c)}\bigr)
  -
  \sum_{j\in R}\operatorname{tr}_{R_j}
  \bigl(t^{(a}t^b t^{c)}\bigr).
\]
If this polynomial is nonzero only for a background field coupled to a global
symmetry, it is an anomaly to be matched rather than an inconsistency.

The representative \(I_4^{(1)}\) is not the invariant datum by itself.  A
local four-dimensional counterterm can move the anomaly among the currents,
or equivalently change the Chern--Simons representative \(I_5^{(0)}\) by a
local transgression.  In an Abelianized chart with several background
one-forms \(A^a\), the elementary Bardeen counterterm
\[
  B_4
  =
  h_{ab,c}\,A^aA^bF^c,
  \qquad h_{ab,c}=-h_{ba,c},
\]
changes a representative written as
\[
  I_4^{(1)}
  =
  d_{abc}\,\lambda^aF^bF^c
\]
to one with
\[
  d'_{abc}
  =
  d_{abc}
  +
  \frac12\bigl(h_{ab,c}+h_{ac,b}\bigr).
\]
The completely symmetric part is unchanged:
\[
  d'_{(abc)}=d_{(abc)}.
\]
This finite algebra is the local form of the statement that counterterms
alter representatives but not the six-form polynomial \(I_6\).  Consequently,
when the next chapter matches the pion Wess--Zumino--Witten functional to the
microscopic flavor anomaly, the coefficient being compared is the symmetric
cohomology coordinate carried by \(I_6\), not an arbitrary distribution of
the anomaly among particular background currents.

\paragraph{Abelianized descent coordinate used in WZW matching.}
\label{par:abelianized-descent-coordinate-wzw}
The coefficient comparison in the Wess--Zumino--Witten application uses a
particularly simple projection of the same relative class.  Restrict the
background connection to a commuting finite-dimensional subspace of the flavor
Lie algebra, so that
\[
  A=A^aT_a,\qquad F^a=\dd A^a,\qquad [T_a,T_b]=0
\]
inside the chosen test subspace.  The degree-three invariant polynomial then
reduces to
\[
  I_6
  =
  C_{abc}\,F^aF^bF^c,
\]
where \(C_{abc}=C_{(abc)}\) is the symmetric coefficient obtained from the
trace in the representation under discussion, multiplied by the normalization
factor fixed above.  On this Abelianized slice a Chern--Simons representative
and its infinitesimal descent are not hidden behind non-Abelian notation:
\begin{equation}
  I_5^{(0)}
  =
  C_{abc}\,A^aF^bF^c,
  \qquad
  \delta_\lambda I_5^{(0)}
  =
  C_{abc}\,\dd\lambda^a F^bF^c
  =
  \dd\!\left(C_{abc}\lambda^aF^bF^c\right),
  \label{eq:abelianized-cubic-descent-coordinate}
\end{equation}
using \(\dd F^a=0\).  Thus the displayed
\(\lambda\,\dd A\,\dd A\) term is not a low-order approximation to a
different anomaly.  It is the Abelianized coordinate of the same descent class
whose full non-Abelian representatives differ by the coboundaries described in
\eqref{eq:relative-anomaly-coboundary}.

For a compact connected semisimple flavor group, the invariant cubic
polynomial is determined by its restriction to a Cartan subalgebra; equivalently,
its symmetric trilinear coefficient is recovered from the homogeneous cubic
function \(P(x)=C_{abc}x^ax^bx^c\) by the polarization identity
\begin{equation}
\begin{aligned}
  C(x,y,z)
  =
  \frac16\bigl[
    &P(x+y+z)-P(x+y)-P(x+z)-P(y+z)
\\
    &\hspace{2.8em}+P(x)+P(y)+P(z)
  \bigr].
\end{aligned}
  \label{eq:cubic-polarization-identity-for-anomaly-coordinate}
\end{equation}
Consequently matching the coefficient of
\(\operatorname{Tr}(\alpha\,\dd A\,\dd A)\) for arbitrary commuting flavor
backgrounds fixes the symmetric cubic anomaly coordinate.  It does not fix a
preferred current representative, and it does not choose a Bardeen convention;
those choices affect \(I_4^{(1)}\) by local counterterm coboundaries while
leaving \(C_{abc}\) unchanged.

\section{Index-Normalized Anomaly Polynomials}
\label{sec:index-normalized-anomaly-polynomials}

The descent representative above is written in the effective-action
normalization convenient for Ward identities.  For global comparisons,
especially when gravitational backgrounds are included, it is useful to keep a
second normalization whose periods are the ones supplied by the Dirac index.
Let \(\nabla^{TM}\) be a Riemannian connection on the tangent bundle with
anti-Hermitian curvature \(\mathsf R\), and let
\(\nabla^E\) be a unitary connection on the fermion bundle \(E_R\) with
anti-Hermitian curvature \(\mathsf F_R\).  Define the Chern--Weil forms
\begin{equation}
  \operatorname{ch}_R(\mathsf F_R)
  =
  \operatorname{tr}_R
  \exp\!\left(\frac{\ii\mathsf F_R}{2\pi}\right),
  \qquad
  p_1(TM)
  =
  -\frac1{8\pi^2}
  \operatorname{tr}_{TM}(\mathsf R^2),
  \label{eq:index-normalized-chern-pontryagin}
\end{equation}
where \(\operatorname{tr}_{TM}\) is the trace in the defining real tangent
representation after complexification.  The six-form anomaly polynomial for a
single left-handed Weyl fermion in \(R\), in the index normalization, is
\begin{equation}
  \mathcal I_6(R)
  =
  \bigl[
    \widehat A(TM)\operatorname{ch}_R(\mathsf F_R)
  \bigr]_6 .
  \label{eq:index-normalized-anomaly-polynomial-general}
\end{equation}
A right-handed Weyl fermion contributes the negative of
\eqref{eq:index-normalized-anomaly-polynomial-general}; equivalently it may be
rewritten as a left-handed fermion in the conjugate representation and then
the same formula may be applied to the left-handed list.

Expanding
\[
  \widehat A(TM)=1-\frac1{24}p_1(TM)+O(\deg 8)
\]
and keeping six-forms gives
\begin{equation}
  \mathcal I_6(R)
  =
  \frac16
  \operatorname{tr}_R
  \left(\frac{\ii\mathsf F_R}{2\pi}\right)^3
  -
  \frac1{24}\,
  p_1(TM)\,
  \operatorname{tr}_R
  \left(\frac{\ii\mathsf F_R}{2\pi}\right).
  \label{eq:four-dimensional-index-polynomial-expanded}
\end{equation}
The first term is the cubic gauge or flavor anomaly.  The second term is the
mixed gauge-gravitational anomaly.  It is present only for symmetries whose
generator has a nonzero trace over the chiral fermion representation.  Thus it
vanishes for a semisimple gauge group in any representation with traceless Lie
algebra generators, but it is a genuine constraint for a gauged \(U(1)\).

The relation between \(\mathcal I_6\) and the effective-action normalization
used in \eqref{eq:consistent-descent-anomaly-representative} is as follows.
Choose a Chern--Simons form \(\mathcal I_5^{(0)}\) with
\[
  d\mathcal I_5^{(0)}=\mathcal I_6,
  \qquad
  \delta_\lambda\mathcal I_5^{(0)}
  =
  d\mathcal I_4^{(1)}(\lambda).
\]
Then the Euclidean anomalous variation of a left-handed determinant is
\begin{equation}
  \delta_\lambda W_E
  =
  2\pi\ii\int_M
  \mathcal I_4^{(1)}(\lambda).
  \label{eq:index-polynomial-effective-action-normalization}
\end{equation}
The anti-Hermitian form \(I_6\) of the previous section is this same descent
data after multiplying by \(2\pi\ii\) and rewriting the connection as
\(\mathsf A=-\ii A\).  Keeping both notations visible prevents a common
coefficient error: the index-normalized polynomial is a characteristic class,
whereas the Ward identity carries the additional factor \(2\pi\ii\) and the
chosen Lorentzian continuation.

\begin{example}[Four-dimensional \(U(1)\) anomaly polynomial]
\label{ex:u1-index-anomaly-polynomial}
Let a left-handed Weyl fermion have integer charge \(q\) under a background
\(U(1)\) connection \(A\) with curvature \(F=dA\).  In the Hermitian curvature
notation \( \ii\mathsf F_R=qF \), and
\eqref{eq:four-dimensional-index-polynomial-expanded} becomes
\begin{equation}
  \mathcal I_6(q)
  =
  \frac{q^3}{6(2\pi)^3}F^3
  -
  \frac{q}{24(2\pi)}p_1(TM)F .
  \label{eq:u1-index-anomaly-polynomial}
\end{equation}
The cubic \(U(1)^3\) coefficient is \(\sum_i q_i^3\) over left-handed Weyl
fields.  The mixed gravitational-\(U(1)\) coefficient is
\(\sum_i q_i\).  If the \(U(1)\) is gauged, both sums must vanish, in addition
to the mixed nonabelian-\(U(1)\) coefficients
\[
  \sum_i q_i\,T_{G_a}(R_i)=0
\]
for each simple nonabelian factor \(G_a\), where
\(\operatorname{tr}_{R_i}(t^a t^b)=T_{G_a}(R_i)\delta^{ab}\) in the chosen
generator normalization.
\end{example}

\begin{example}[\(SU(N)\) cubic anomaly in fundamental coordinates]
\label{ex:su-n-fundamental-adjoint-anomaly}
Let \(G=SU(N)\), \(N\geq3\), and choose a reference symmetric tensor
\[
  d_\square^{abc}
  =
  \operatorname{tr}_{\square}
  \bigl(t^{(a}t^b t^{c)}\bigr)
\]
in the monograph trace convention.  For an irreducible representation \(R\)
define \(A(R)\) by
\[
  \operatorname{tr}_{R}
  \bigl(t^{(a}t^b t^{c)}\bigr)
  =
  A(R)d_\square^{abc}.
\]
This equation defines the cubic-anomaly coordinate; rescaling all generators
rescales both sides and leaves \(A(R)\) unchanged.  The conjugate
representation obeys \(A(\overline R)=-A(R)\), and every real representation
has \(A(R)=0\).  In particular,
\[
  A(\square)=1,
  \qquad
  A(\overline\square)=-1,
  \qquad
  A(\operatorname{adj})=0 .
\]
Therefore a chiral \(SU(N)\) gauge theory containing
\(n_\square\) left-handed fundamentals,
\(n_{\overline\square}\) left-handed antifundamentals, and any number of
left-handed adjoints has cubic gauge anomaly
\begin{equation}
  d_{\rm total}^{abc}
  =
  (n_\square-n_{\overline\square})d_\square^{abc}.
  \label{eq:su-n-fund-anti-adjoint-anomaly}
\end{equation}
A vectorlike Dirac fundamental consists, in a left-handed description, of one
\(\square\) and one \(\overline\square\); its contribution to
\eqref{eq:su-n-fund-anti-adjoint-anomaly} vanishes.
\end{example}

\begin{example}[One Standard Model generation]
\label{ex:standard-model-anomaly-sums}
Use left-handed Weyl fields
\[
  Q_L:(3,2)_{1/6},\qquad
  u^c:(\overline3,1)_{-2/3},\qquad
  d^c:(\overline3,1)_{1/3},
\]
\[
  L_L:(1,2)_{-1/2},\qquad
  e^c:(1,1)_1 .
\]
The mixed nonabelian-\(U(1)_Y\) sums, with the common quadratic index of the
fundamental factored out, are
\[
  SU(3)^2U(1)_Y:
  \quad
  2\!\left(\frac16\right)
  -\frac23
  +\frac13
  =
  0,
\]
and
\[
  SU(2)^2U(1)_Y:
  \quad
  3\!\left(\frac16\right)
  -\frac12
  =
  0.
\]
The cubic and mixed gravitational hypercharge sums are
\begin{equation}
\begin{aligned}
  \sum_{\rm Weyl}Y^3
  &=
  6\left(\frac16\right)^3
  +3\left(-\frac23\right)^3
  +3\left(\frac13\right)^3
  +2\left(-\frac12\right)^3
  +1
  =
  0,
\\
  \sum_{\rm Weyl}Y
  &=
  6\left(\frac16\right)
  +3\left(-\frac23\right)
  +3\left(\frac13\right)
  +2\left(-\frac12\right)
  +1
  =
  0 .
\end{aligned}
  \label{eq:standard-model-hypercharge-anomaly-sums}
\end{equation}
The same field list contains \(3+1=4\) weak doublets, so the finite
\(SU(2)\) anomaly discussed in Chapter~\ref{ch:volume-ii-21} is absent for
each generation.  These cancellations are arithmetic consequences of the
charge assignments; they do not follow from perturbative renormalizability
alone.
\end{example}

\begin{remark}[Anomaly matching as equality of anomaly lines]
\label{rem:anomaly-matching-as-line-equality}
Consider an RG flow from an ultraviolet theory \(\mathcal Q_{\rm UV}\) to an
infrared description \(\mathcal Q_{\rm IR}\), and suppose a background-field
groupoid \(\mathcal B(M)\) for a global symmetry is defined along the flow.
Assume first a finite regulator.  For a cutoff \(\Lambda\), the regulated
partition vector is a section of an anomaly line over \(\mathcal B(M)\); in a
local frame its background-gauge transport is the finite cochain
\(C_\Lambda(g;B)\).  A Wilsonian step
\(\Lambda\to\Lambda'\) is a pushforward over the shell variables at fixed
retained fields and fixed background.  If the regulator split, shell domain,
and shell measure are equivariant for the background action, the change of
variables in the regulated integral gives
\[
  R_{\Lambda\to\Lambda'}\bigl(T_g Z_\Lambda\bigr)
  =
  T_g\,R_{\Lambda\to\Lambda'}(Z_\Lambda),
\]
where \(T_g\) is the anomaly-line transport.  In a chosen frame this is the
same finite change-of-variables identity as
\eqref{eq:finite-regulator-shell-inherits-anomaly}: the measure cocycle is
inherited by the Wilsonian section, while a change of local regulator or
retained-field frame adds only a coboundary counterterm.

The locality input is narrower than the matching statement.  Quasi-locality of
the shell integration is used to ensure that differences between nearby
regulator prescriptions, shell cuts, and subtraction frames are local
functionals of the retained fields and backgrounds.  Their background
variation is therefore a local counterterm coboundary.  This does not assert
that a full one-particle-irreducible functional, or a final endpoint partition
function after massless fields have been integrated, differs from the
ultraviolet object by a local background-only factor.

Taking the continuum and infrared limits is a further assumption: the proposed
infrared description must define a section over the same background groupoid
and the same anomaly line for the exact symmetry being matched.  A phase with
spontaneous symmetry breaking realizes that section through Goldstone fields
and a gauged Wess--Zumino--Witten term, possibly tensored with additional
massless or topological sectors.  A purely topological infrared response is an
invertible boundary condition for the same anomaly theory.  This is the local
descent version of the anomaly-line and inflow language in
Chapter~\ref{ch:global-anomaly-inflow-invertible-field-theories}, and the
determinant/Pfaffian-line holonomy version in
Chapter~\ref{ch:curved-eta-invariants-global-anomalies}.  The statement is
therefore a conditional structural principle: under these finite-regulator and
endpoint hypotheses, UV and IR anomaly classes agree modulo local
counterterms, but the infrared realization need not be unique.
\end{remark}

\section{Anomaly Inflow for Local Anomalies}

The descent equations have a second interpretation when the four-dimensional
theory is placed on the boundary of a five-dimensional invertible response.
Let \(X\) be an oriented five-manifold with boundary
\(\partial X=M\), and let \(\mathsf A_X\) be a smooth background connection on
\(X\) whose restriction to \(M\) is the four-dimensional connection
\(\mathsf A\).  The local Chern--Simons functional
\begin{equation}
  W_X^{\mathrm{bulk}}[\mathsf A_X]
  =
  -\int_X I_5^{(0)}(\mathsf A_X)
  \label{eq:local-inflow-bulk-functional}
\end{equation}
has gauge variation
\begin{equation}
  \delta_\lambda W_X^{\mathrm{bulk}}
  =
  -\int_X \dd I_4^{(1)}(\lambda,\mathsf A_X)
  =
  -\int_M I_4^{(1)}(\lambda,\mathsf A).
  \label{eq:local-inflow-bulk-variation}
\end{equation}
Writing the four-dimensional effective action in
\eqref{eq:consistent-descent-anomaly-representative} as \(W_M[\mathsf A]\),
one obtains the bulk-boundary identity
\begin{equation}
  \delta_\lambda
  \left(
    W_M[\mathsf A]+W_X^{\mathrm{bulk}}[\mathsf A_X]
  \right)
  =
  \int_M I_4^{(1)}(\lambda,\mathsf A)
  -
  \int_{\partial X} I_4^{(1)}(\lambda,\mathsf A_X)
  =
  0,
  \label{eq:local-anomaly-inflow-cancellation}
\end{equation}
where the boundary orientation on \(\partial X\) is the orientation of \(M\).
This is anomaly inflow.  The boundary theory by itself transforms by the
consistent anomaly; the bulk Chern--Simons response transforms by the opposite
boundary term.  The gauge-invariant object is the combined bulk-boundary
system.

In the Callan--Harvey realization, \(X\) is a physical gapped bulk or a
domain-wall geometry, and \(M\) supports chiral modes localized at the wall.
Integrating out the massive bulk modes produces a Chern--Simons response whose
coefficient jumps across the wall.  The jump supplies precisely the term
\(-\int_M I_4^{(1)}\) in \eqref{eq:local-inflow-bulk-variation}, while the
localized chiral modes supply \(+\int_M I_4^{(1)}\).  The boundary chiral
modes retain their consistent anomaly, and gauge invariance is a property of
the combined bulk-boundary system.

The formula \eqref{eq:local-inflow-bulk-functional} is local in a chosen
trivialization.  On topologically nontrivial bundles the Chern--Simons form is
defined patchwise, and the exponentiated bulk functional must be defined by
differential-geometric data with integral periods, equivalently by an
invertible five-dimensional field theory whose curvature is the anomaly
polynomial \(I_6\).  The local descent computation fixes the infinitesimal
variation.  Global questions require the finite version: a closed
five-manifold carrying a background field is assigned a phase, and cutting it
open gives the anomalous transformation law of a boundary theory.  That finite
form is the starting point of the global-anomaly discussion in
Chapter~\ref{ch:volume-ii-21}.

\section{Consistent and Covariant Currents}

The current obtained by differentiating the effective action,
\[
  J_{\mathrm{cons}}^{a\mu}(x)
  =
  \frac{\delta W[A]}{\delta A^a_\mu(x)},
\]
is the consistent current.  Its anomaly is the variation of a functional and
therefore obeys the Wess--Zumino consistency condition.  This integrability is
the reason that the consistent anomaly is the object classified by
\(H^{1,4}(s\mid d)\).  More precisely, the anomaly class modulo local
counterterms is the consistent descent class represented by
\(\int_M I_4^{(1)}\); covariant Ward identities are obtained after taking a
local differential of this class and adding a Bardeen--Zumino polynomial.

The consistent current need not transform covariantly under background gauge
transformations.  There is a local polynomial current
\[
  J_{\mathrm{BZ}}^{a\mu}(A)
\]
called the Bardeen--Zumino current, whose addition gives a covariant current
\[
  J_{\mathrm{cov}}^{a\mu}
  =
  J_{\mathrm{cons}}^{a\mu}+J_{\mathrm{BZ}}^{a\mu}.
\]
In the anti-Hermitian form convention used in the descent formula above, the
dual three-form of the Bardeen--Zumino current for one left-handed fermion in
the representation \(R\) may be chosen as
\begin{equation}
  \star J_{\mathrm{BZ}}^a
  =
  \kappa\,
  \operatorname{tr}_R
  \left[
    T_R^a
    \left(
      \mathsf A\mathsf F+\mathsf F\mathsf A
      -\frac12\mathsf A^3
    \right)
  \right],
  \qquad
  \kappa=\frac{\ii}{24\pi^2},
  \label{eq:bardeen-zumino-current-polynomial}
\end{equation}
where \(T_R^a\) is the anti-Hermitian generator used to vary
\(\mathsf A\), products include wedge product and matrix multiplication, and
\(\star J_{\mathrm{BZ}}^a\) denotes
\[
  \star J_{\mathrm{BZ}}^a
  =
  \frac1{3!}
  \epsilon_{\mu\nu\rho\sigma}
  J_{\mathrm{BZ}}^{a\mu}
  \dd x^\nu\wedge\dd x^\rho\wedge\dd x^\sigma .
\]
Equivalently, with
\[
  \mathsf A=\mathsf A_\mu\dd x^\mu,\qquad
  \mathsf F=\frac12\mathsf F_{\rho\sigma}\dd x^\rho\wedge\dd x^\sigma,
\]
the component current is the local polynomial
\begin{equation}
  J_{\mathrm{BZ}}^{a\mu}
  =
  \frac{\kappa}{2}\,
  \epsilon^{\mu\nu\rho\sigma}
  \operatorname{tr}_R
  \left[
    T_R^a
    \left(
      \mathsf A_\nu\mathsf F_{\rho\sigma}
      +
      \mathsf F_{\rho\sigma}\mathsf A_\nu
      -
      \mathsf A_\nu\mathsf A_\rho\mathsf A_\sigma
    \right)
  \right].
  \label{eq:bardeen-zumino-current-component-curvature}
\end{equation}
Expanding
\[
  \mathsf F_{\rho\sigma}
  =
  \partial_\rho\mathsf A_\sigma
  -
  \partial_\sigma\mathsf A_\rho
  +
  [\mathsf A_\rho,\mathsf A_\sigma]
\]
and using antisymmetry of
\(\epsilon^{\mu\nu\rho\sigma}\) gives the derivative form
\begin{equation}
  J_{\mathrm{BZ}}^{a\mu}
  =
  \kappa\,\epsilon^{\mu\nu\rho\sigma}
  \operatorname{tr}_R
  \left[
    T_R^a
    \left(
      \mathsf A_\nu\partial_\rho\mathsf A_\sigma
      +
      \partial_\rho\mathsf A_\sigma\,\mathsf A_\nu
      +
      \frac12\mathsf A_\nu\mathsf A_\rho\mathsf A_\sigma
      +
      \mathsf A_\rho\mathsf A_\sigma\mathsf A_\nu
    \right)
  \right].
  \label{eq:bardeen-zumino-current-component-derivative}
\end{equation}
The two displayed component expressions are the same polynomial written once
with \(\mathsf F\) and once with \(\partial\mathsf A\).  The ordering of
matrices in \eqref{eq:bardeen-zumino-current-component-derivative} is part of
the convention; replacing it by a commutative shorthand would lose the
distinction between consistent and covariant currents for a nonabelian
background.

\paragraph{Abelian calibration of the Bardeen--Zumino factor.}
The simplest check of the normalization is the one-generator commuting
subspace.  Let
\[
  I_6=C\,\mathsf F^3,\qquad \mathsf F=\dd\mathsf A ,
\]
where \(C\) includes the trace and the factor \(\kappa\).  The corresponding
Chern--Simons and consistent descent forms are
\begin{equation}
  I_5^{(0)}=C\,\mathsf A\mathsf F^2,\qquad
  I_4^{(1)}(\lambda,\mathsf A)=C\,\lambda\,\mathsf F^2 .
  \label{eq:abelian-consistent-descent-calibration}
\end{equation}
Thus the consistent Ward identity carries the coefficient \(C\).  In the
same Abelianized coordinate the Bardeen--Zumino current
\eqref{eq:bardeen-zumino-current-polynomial} reduces to
\begin{equation}
  \star J_{\mathrm{BZ}}
  =
  2C\,\mathsf A\mathsf F ,
  \qquad
  \dd\star J_{\mathrm{BZ}}
  =
  2C\,\mathsf F^2 ,
  \label{eq:abelian-bz-divergence-calibration}
\end{equation}
because \(\mathsf A^3=0\) for a single one-form and \(\dd\mathsf F=0\).
The covariant Ward representative obtained by improving the current is
therefore
\begin{equation}
  C\,\lambda\,\mathsf F^2
  +
  2C\,\lambda\,\mathsf F^2
  =
  3C\,\lambda\,\mathsf F^2 .
  \label{eq:abelian-covariant-anomaly-factor-three}
\end{equation}
This factor of three is a normalization check, not a new anomaly class.  The
first term is the descent representative of the effective-action variation;
the additional two units are the local divergence of the Bardeen--Zumino
polynomial.  In the nonabelian case the same statement is encoded by the
ordered polynomial in
\eqref{eq:bardeen-zumino-current-component-derivative}; the covariant current
transforms covariantly, but it is not in general the differential of a local
effective action.
For a set of Weyl fermions, sum \eqref{eq:bardeen-zumino-current-polynomial}
over left-handed representations and subtract the corresponding right-handed
representations, in parallel with the anomaly polynomial \(I_6\).
The divergence of \(J_{\mathrm{cov}}\) is gauge covariant and is often the most
compact local expression of the anomaly.  However,
\(J_{\mathrm{cov}}\) is generally not the functional derivative of a local
effective action.  Thus it is not the current that appears directly in the
integrability condition \eqref{eq:wess-zumino-consistency}.  The consistent
anomaly is the cohomology class.  The covariant anomaly is obtained from a
particular representative by applying the local differential of that class and
adding the Bardeen--Zumino polynomial; it is a covariant representative of the
Ward identity, not a representative of a second obstruction class.
Gauge-anomaly cancellation is therefore a statement about the consistent
class; after that class vanishes, covariant currents may be used as convenient
representatives of gauge-covariant Ward identities.

\section{Local Counterterms and Preserved Currents}

An anomaly representative depends on the choice of local counterterms, while
its cohomology class does not.  Let \(B\) denote background gauge fields for a
set of currents, and let \(W[B]\) be the renormalized generating functional.
Adding a local functional \(I_{\rm ct}[B]\) changes
\[
  W[B]\longmapsto W[B]+I_{\rm ct}[B].
\]
If a background gauge transformation with parameter \(\zeta\) has anomaly
\[
  \delta_\zeta W[B]=\mathcal A(\zeta,B),
\]
then after the counterterm shift
\[
  \mathcal A(\zeta,B)
  \longmapsto
  \mathcal A(\zeta,B)+\delta_\zeta I_{\rm ct}[B].
\]
The second term is BRST-exact in the local functional complex.  It can move
the displayed divergence among currents that are coupled to backgrounds, but
it cannot remove a nonzero anomaly class.

This distinction is essential when one of the currents is gauged dynamically.
The gauge current must satisfy the Slavnov--Taylor identity; otherwise the
gauge-fixed state space cannot be quotiented to a positive physical Hilbert
space.  For a vector and an axial current coupled to background fields, a
Bardeen counterterm can shift the local expression of the anomaly between
vector and axial Ward identities.  The convention used in
\eqref{eq:abelian-axial-anomaly} is the one in which the vector gauge Ward
identity is preserved and the axial current carries the anomaly.  For a
purely global symmetry, by contrast, the anomaly is allowed and becomes an
RG-invariant background response to be matched by the infrared theory.

\section{The Strong CP Parameter}

In QCD with \(N_f\) quark flavors, the Euclidean gauge action may include the
topological term
\[
  \ii\theta\,
  \frac1{32\pi^2}
  \int
  \epsilon^{\mu\nu\rho\sigma}
  \operatorname{tr}(F_{\mu\nu}F_{\rho\sigma})\,\dd^4x .
\]
The complex quark mass matrix \(M\) was introduced in
\eqref{eq:qcd-chiral-mass-lagrangian}.  A flavor-singlet axial rotation
changes the phase of \(\det M\) and, by the anomaly, shifts \(\theta\).  With
the convention used in this chapter, the invariant CP-odd parameter is
\begin{equation}
  \bar\theta
  =
  \theta+\arg\det M .
  \label{eq:qcd-physical-theta}
\end{equation}
If the sign convention for \(\gamma_5\) or for the topological density is
reversed, both transformation laws change together and the invariant
combination is correspondingly rewritten.  The invariant statement is that
only the anomaly-invariant combination of the topological angle and the quark
mass phase is physically meaningful.

\section{Topological Charge Sectors and Periodicity}

For gauge group \(SU(N_c)\) on a compact Euclidean four-manifold, or on
\(\mathbb R^4\) with finite-action boundary conditions so that the connection
extends to a bundle over \(S^4\), the topological charge is
\[
  Q[A]
  =
  \frac1{32\pi^2}
  \int
  \epsilon^{\mu\nu\rho\sigma}
  \operatorname{tr}(F_{\mu\nu}F_{\rho\sigma})\,\dd^4x .
\]
With \(F\) regarded as the matrix-valued curvature of the fundamental
connection and with the matrix trace convention fixed in
Chapter~\ref{ch:vol2-classical-yang-mills-matter}, \(Q[A]\in\mathbb Z\).
This integer statement is a Chern--Weil statement about the curvature form;
rewriting it in components must use the same generator normalization as the
one used to define \(F=A^a t^a\).  The Euclidean functional integral
decomposes formally into connected components of fixed topological charge:
\[
  Z(\theta)
  =
  \sum_{Q\in\mathbb Z}
  \ee^{\ii\theta Q}Z_Q .
\]
This expression makes the \(2\pi\)-periodicity in \(\theta\) manifest when no
additional global identifications modify the allowed bundles or background
fields.

The axial anomaly explains why \(\theta\) cannot be assigned independently of
the fermion mass phases.  A flavor-singlet chiral rotation changes the
fermion measure by a phase proportional to \(Q[A]\) and changes
\(\arg\det M\) by the opposite amount in the mass term.  The sector
decomposition therefore gives the same invariant parameter
\(\bar\theta=\theta+\arg\det M\) as the local anomalous Ward identity.  The
local identity and the topological decomposition are two presentations of the
same anomalous Jacobian in a gauge background with nonzero topological charge.

\section{Topological Susceptibility as a Regulated Cumulant}
\label{sec:topological-susceptibility-regulated-cumulant}

The phrase ``topological susceptibility'' is often used as if it were a
number attached directly to a formal continuum expression
\(\langle Q^2\rangle/V_4\).  The object needed in QCD is more precise: it is a
thermodynamic and continuum limit of a cumulant in a specified regulator.

\begin{definition}[Finite-regulator theta data]
\label{def:finite-regulator-theta-data}
Fix a gauge-invariant Euclidean regulator on an oriented four-volume \(V_4\).
Finite-regulator theta data consist of nonnegative sector weights
\(Z_{\Lambda,V_4,Q}\), \(Q\in\mathbb Z\), for the vectorlike theory at
\(\theta=0\), such that
\begin{equation}
  Z_{\Lambda,V_4}(\theta)
  =
  \sum_{Q\in\mathbb Z}
  \ee^{\ii\theta Q}Z_{\Lambda,V_4,Q}
  \label{eq:finite-regulator-theta-partition-function}
\end{equation}
is finite for \(\theta\) in a neighborhood of zero.  The finite-regulator
vacuum-energy density near \(\theta=0\) is
\begin{equation}
  E_{\Lambda,V_4}(\theta)
  =
  -\frac1{V_4}\log Z_{\Lambda,V_4}(\theta),
  \label{eq:finite-regulator-theta-energy}
\end{equation}
where the logarithm is the branch continuous from \(\theta=0\).
\end{definition}

\begin{remark}[Finite-volume susceptibility identity]
\label{rem:finite-volume-topological-susceptibility}
For every finite-regulator theta data set,
\begin{equation}
  E_{\Lambda,V_4}''(0)
  =
  \frac1{V_4}
  \left(
    \langle Q^2\rangle_{\Lambda,V_4,0}
    -
    \langle Q\rangle_{\Lambda,V_4,0}^2
  \right),
  \label{eq:finite-volume-topological-susceptibility}
\end{equation}
where the expectation is with respect to the positive weights
\(Z_{\Lambda,V_4,Q}/Z_{\Lambda,V_4}(0)\).  If CP is preserved at
\(\theta=0\), then \(\langle Q\rangle_{\Lambda,V_4,0}=0\), and the right hand
side is \(V_4^{-1}\langle Q^2\rangle_{\Lambda,V_4,0}\).
Differentiate \eqref{eq:finite-regulator-theta-partition-function}:
\[
  Z'(0)=\ii\sum_Q Q Z_Q,
  \qquad
  Z''(0)=-\sum_Q Q^2Z_Q ,
\]
where the regulator and volume labels are suppressed.  Hence
\[
  \frac{\dd^2}{\dd\theta^2}\log Z(\theta)\bigg|_{\theta=0}
  =
  \frac{Z''(0)}{Z(0)}
  -
  \left(\frac{Z'(0)}{Z(0)}\right)^2
  =
  -\langle Q^2\rangle+\langle Q\rangle^2 .
  \]
Multiplying by \(-1/V_4\) gives
\eqref{eq:finite-volume-topological-susceptibility}.  CP sends
\(Q\mapsto -Q\); if the \(\theta=0\) weights are CP invariant, the first
moment vanishes.
\end{remark}

\begin{proposition}[Finite-regulator theta-cumulant hierarchy]
\label{prop:finite-regulator-theta-cumulant-hierarchy}
For finite-regulator theta data, define the charge cumulants at \(\theta=0\)
by
\begin{equation}
  \mathcal K_Q(t)
  =
  \log\left\langle \ee^{tQ}\right\rangle_{\Lambda,V_4,0}
  =
  \sum_{n\ge1}\kappa_n^{(Q)}\frac{t^n}{n!}.
  \label{eq:finite-regulator-charge-cumulant-generator}
\end{equation}
Then, in the analytic neighborhood where
Definition~\ref{def:finite-regulator-theta-data} applies,
\begin{equation}
  E_{\Lambda,V_4}(\theta)-E_{\Lambda,V_4}(0)
  =
  -\frac1{V_4}\,\mathcal K_Q(\ii\theta),
  \qquad
  \partial_\theta^n E_{\Lambda,V_4}(0)
  =
  -\frac{\ii^n}{V_4}\,\kappa_n^{(Q)} .
  \label{eq:finite-regulator-theta-cumulant-derivatives}
\end{equation}
If the \(\theta=0\) weights are CP invariant, all odd charge cumulants vanish
and the finite-volume expansion can be written
\begin{equation}
  E_{\Lambda,V_4}(\theta)-E_{\Lambda,V_4}(0)
  =
  \frac12\chi_{\Lambda,V_4}\theta^2
  \left(
    1+b_{2,\Lambda,V_4}\theta^2
     +b_{4,\Lambda,V_4}\theta^4+O(\theta^6)
  \right),
  \label{eq:finite-regulator-theta-even-expansion}
\end{equation}
with
\begin{equation}
  \chi_{\Lambda,V_4}
  =
  \frac{\kappa_2^{(Q)}}{V_4},\qquad
  b_{2,\Lambda,V_4}
  =
  -\frac{\kappa_4^{(Q)}}{12\kappa_2^{(Q)}},\qquad
  b_{4,\Lambda,V_4}
  =
  \frac{\kappa_6^{(Q)}}{360\kappa_2^{(Q)}} ,
  \label{eq:finite-regulator-theta-b2-b4-cumulants}
\end{equation}
provided \(\kappa_2^{(Q)}\neq0\).

These are finite-volume identities, not yet continuum statements.  In the
thermodynamic limit the derivatives in
\eqref{eq:finite-regulator-theta-cumulant-derivatives} must be taken on a
selected analytic branch of the vacuum energy; branch crossings or
spontaneous CP breaking can obstruct a single Taylor series.  In addition, if
the renormalization convention permits adding a local vacuum counterterm
\(f_{\rm ct}(\theta)\), then the displayed derivatives shift by
\(f_{\rm ct}^{(n)}(0)\).  The same counterterm convention must therefore be
used for \(\chi_{\rm top}\), \(b_2\), \(b_4\), and any Witten--Veneziano
matching data.
\end{proposition}

\begin{proof}
Divide \eqref{eq:finite-regulator-theta-partition-function} by
\(Z_{\Lambda,V_4}(0)\).  The ratio is the characteristic function
\(\langle\ee^{\ii\theta Q}\rangle_{\Lambda,V_4,0}\), hence
\[
  \log\frac{Z_{\Lambda,V_4}(\theta)}{Z_{\Lambda,V_4}(0)}
  =
  \mathcal K_Q(\ii\theta).
\]
Subtracting \(E_{\Lambda,V_4}(0)\) from
\eqref{eq:finite-regulator-theta-energy} gives the first identity in
\eqref{eq:finite-regulator-theta-cumulant-derivatives}; differentiating gives
the second.  CP invariance pairs the \(Q\) and \(-Q\) sector weights, so
\(\mathcal K_Q(t)=\mathcal K_Q(-t)\) and the odd cumulants vanish.  Expanding
the even part gives
\[
  E(\theta)-E(0)
  =
  \frac{\kappa_2^{(Q)}}{2V_4}\theta^2
  -
  \frac{\kappa_4^{(Q)}}{24V_4}\theta^4
  +
  \frac{\kappa_6^{(Q)}}{720V_4}\theta^6+\cdots .
\]
Comparison with \eqref{eq:finite-regulator-theta-even-expansion} gives
\eqref{eq:finite-regulator-theta-b2-b4-cumulants}.  The last paragraph is the
definition of branchwise differentiability and local counterterm covariance
for the finite-volume identities just derived.
\end{proof}

\begin{proposition}[Local density form and contact convention]
\label{prop:finite-regulator-susceptibility-density-cumulant}
Assume that the finite regulator is placed on a closed four-volume, such as a
periodic box, or on a volume with boundary data for which the topological
charge is fixed, and that it supplies a regulated local density
\(q_{\Lambda,V_4}(x)\) satisfying
\begin{equation}
  Q
  =
  \int_{V_4} q_{\Lambda,V_4}(x)\,\dd^4x
  \label{eq:finite-regulator-local-topological-charge}
\end{equation}
configuration by configuration.  Then the finite-volume susceptibility is
equivalently, with the integrals interpreted as the regulator's finite
sum or integral,
\begin{equation}
  E_{\Lambda,V_4}''(0)
  =
  \frac1{V_4}
  \int_{V_4}\dd^4x
  \int_{V_4}\dd^4y\,
  \bigl\langle
    q_{\Lambda,V_4}(x)q_{\Lambda,V_4}(y)
  \bigr\rangle_{\Lambda,V_4,0}^{\rm c}.
  \label{eq:finite-regulator-density-susceptibility}
\end{equation}
If the regulator and state are translation invariant on a periodic volume,
this reduces to the zero-momentum two-point function
\begin{equation}
  E_{\Lambda,V_4}''(0)
  =
  \int_{V_4}\dd^4x\,
  \bigl\langle
    q_{\Lambda,V_4}(x)q_{\Lambda,V_4}(0)
  \bigr\rangle_{\Lambda,V_4,0}^{\rm c}.
  \label{eq:periodic-density-susceptibility}
\end{equation}
\end{proposition}

\begin{proof}
Insert \eqref{eq:finite-regulator-local-topological-charge} into the finite
cumulant:
\[
  \langle Q^2\rangle-\langle Q\rangle^2
  =
  \int_{V_4}\dd^4x
  \int_{V_4}\dd^4y\,
  \left(
    \langle q_{\Lambda,V_4}(x)q_{\Lambda,V_4}(y)\rangle
    -
    \langle q_{\Lambda,V_4}(x)\rangle
    \langle q_{\Lambda,V_4}(y)\rangle
  \right).
\]
Dividing by \(V_4\) and using
\eqref{eq:finite-volume-topological-susceptibility} gives
\eqref{eq:finite-regulator-density-susceptibility}.  Translation invariance
sets the connected correlator equal to a function of \(x-y\); the double
integral is then \(V_4\) times the integral over one relative coordinate.
\end{proof}

The continuum formula
\(\chi_{\rm top}=\int\dd^4x\,\langle q(x)q(0)\rangle^{\rm c}\) is therefore
not a bare multiplication rule for two formal \(F\widetilde F\) densities at
coincident points.  It is the limit of the regulated zero-momentum curvature
above, with the contact term selected by the same local \(\theta\)-counterterm
convention used in
Definition~\ref{def:topological-susceptibility}.  Changing the convention by
an additive local vacuum-energy term \(\frac12 c_\theta\theta^2\) shifts both
\(E''(0)\) and the integrated contact part of the density two-point function
by \(c_\theta\).  Thus Witten--Veneziano matching uses the renormalized
zero-momentum curvature of pure Yang--Mills, not an unrenormalized
point-split product with its contact term silently discarded.

\begin{definition}[Topological susceptibility]
\label{def:topological-susceptibility}
The topological susceptibility of a continuum theory in a specified phase and
regulator class is the limit
\begin{equation}
  \chi_{\rm top}
  =
  \lim_{\Lambda\to\infty}\lim_{V_4\to\infty}
  E_{\Lambda,V_4}''(0),
  \label{eq:topological-susceptibility-limit}
\end{equation}
when this limit exists and is independent of the allowed regulator choices
after the local counterterm convention for the theta term has been fixed.
For pure Yang--Mills we write the corresponding quantity as
\(\chi_{\rm YM}\).
\end{definition}

This definition intentionally separates three facts which are often conflated.
The finite identity
\eqref{eq:finite-volume-topological-susceptibility} is elementary.  The
existence and regulator independence of
\eqref{eq:topological-susceptibility-limit} is a nonperturbative statement
about the theory.  The use of \(\chi_{\rm YM}\) inside a chiral effective
description with dynamical quarks is a further large-\(N_c\) matching
assumption, developed in the next chapter.

\begin{controlledapproximation}[Fixed-topology extraction and bias]
\label{ca:fixed-topology-theta-extraction}
The sector weights themselves carry physical information, but a simulation or
semiclassical ensemble constrained to one integer \(Q\) is not the
\(\theta=0\) vacuum.  Let a selected CP-even branch have, in a neighborhood of
\(\theta=0\),
\begin{equation}
  E(\theta)-E(0)
  =
  \frac12\chi\theta^2
  \left(1+b_2\theta^2+R_6(\theta)\right),
  \qquad \chi>0 ,
  \label{eq:theta-fixed-topology-branch-expansion}
\end{equation}
and assume \(\chi V_4\gg1\), no competing branch on the saddle contour, and
controlled endpoint and \(R_6\) remainders.  Fourier inversion of the sector
sum gives the fixed-topology weight
\begin{equation}
  \widehat Z_Q
  =
  \int_{-\pi}^{\pi}\frac{\dd\theta}{2\pi}\,
  \exp\!\left[-V_4E(\theta)-\ii Q\theta\right].
  \label{eq:fixed-topology-fourier-inversion}
\end{equation}
Keeping the Gaussian saddle and the first quartic correction gives
\begin{equation}
  \frac{\widehat Z_Q}{\ee^{-V_4E(0)}}
  =
  \frac{1}{\sqrt{2\pi\chi V_4}}\,
  \exp\!\left[-\frac{Q^2}{2\chi V_4}\right]
  \left[
    1-\frac{b_2}{2}
    \left(
      \frac{3}{\chi V_4}
      -\frac{6Q^2}{(\chi V_4)^2}
      +\frac{Q^4}{(\chi V_4)^3}
    \right)
    +R_Q
  \right],
  \label{eq:fixed-topology-gaussian-sector-weight}
\end{equation}
where \(R_Q\) contains the sixth-and-higher theta cumulants, endpoint tails,
and branch-selection defects in the declared \(Q^2/(\chi V_4)\) window.
Thus the curvature \(\chi\) can be extracted from the Gaussian slope of
\(\log\widehat Z_Q\) as a function of \(Q^2\), but not from the variance of
\(Q\) inside a single fixed sector.  In a fixed-\(Q\) ensemble the random
variable \(Q\) has zero variance by construction.

The same saddle controls local observables.  If a CP-even local observable has
branch expectation
\[
  O(\theta)=O_0+\frac12 O_2\theta^2+O(\theta^4),
\]
then its fixed-topology expectation is
\begin{equation}
  O_Q
  =
  O_0
  +
  \frac{O_2}{2\chi V_4}
  \left(
    1-\frac{Q^2}{\chi V_4}
  \right)
  +R_{O,Q}.
  \label{eq:fixed-topology-local-observable-bias}
\end{equation}
The \(1/V_4\) correction is the leading fixed-topology bias.  It is harmless
only when it is smaller than the stated statistical, continuum, and
finite-volume error budget.  Otherwise fixed-sector measurements must either
fit the sector weights \(\widehat Z_Q\), reweight in \(\theta\), or subtract
\eqref{eq:fixed-topology-local-observable-bias} with independently controlled
\(\chi\) and \(O_2\).  This is why topological freezing is an observable
systematic, beyond an algorithmic autocorrelation problem.
\end{controlledapproximation}

\section{Theta Branches and Cluster Decomposition}
\label{sec:theta-branches-cluster-decomposition}

The sector sum
\(\sum_Q\ee^{\ii\theta Q}Z_Q\) explains \(2\pi\)-periodicity, but it does not
by itself identify a pure infinite-volume vacuum.  In a theory with several
large-volume branches the thermodynamic limit selects the branch of smallest
energy density, except at degeneracy loci where a phase-selection prescription
is needed.  Cluster decomposition is the local diagnostic that one has chosen
a pure phase rather than a convex mixture.

\begin{definition}[Theta-branch data]
\label{def:theta-branch-data}
Theta-branch data for a regulated Euclidean gauge theory in a volume \(V_4\)
consists of a finite or countable branch set \(\mathcal B\), branch partition
functions \(Z_{\Lambda,V_4,\beta}(\theta)\), and local normalized states
\(\omega_{\Lambda,V_4,\beta,\theta}\) on gauge-invariant observables, such
that
\begin{equation}
  Z_{\Lambda,V_4}(\theta)
  =
  \sum_{\beta\in\mathcal B}Z_{\Lambda,V_4,\beta}(\theta),
  \qquad
  p_{\Lambda,V_4,\beta}(\theta)
  =
  \frac{Z_{\Lambda,V_4,\beta}(\theta)}
       {Z_{\Lambda,V_4}(\theta)} .
  \label{eq:theta-branch-partition-decomposition}
\end{equation}
When the limits exist, write
\begin{equation}
  E_\beta(\theta)
  =
  -\lim_{\Lambda\to\infty}\lim_{V_4\to\infty}
  \frac1{V_4}\log Z_{\Lambda,V_4,\beta}(\theta)
  \label{eq:theta-branch-energy-density}
\end{equation}
for the branch energy density.  The data are \(2\pi\)-periodic by branch
relabeling when there is a bijection \(T:\mathcal B\to\mathcal B\) with
\begin{equation}
  E_{T\beta}(\theta+2\pi)=E_\beta(\theta).
  \label{eq:theta-branch-periodic-relabeling}
\end{equation}
\end{definition}

\begin{remark}[Thermodynamic branch selection]
\label{rem:theta-thermodynamic-branch-selection}
Assume a finite branch set and a fixed \(\theta\) at which the limiting branch
energies \(E_\beta(\theta)\) exist.  If a unique branch \(\beta_*\) satisfies
\begin{equation}
  E_{\beta_*}(\theta)
  <
  E_\beta(\theta)
  \qquad
  \hbox{for every }\beta\neq\beta_*,
  \label{eq:unique-theta-branch-minimum}
\end{equation}
and if each branch partition function has the form
\begin{equation}
  Z_{\Lambda,V_4,\beta}(\theta)
  =
  \exp\{-V_4E_\beta(\theta)+o(V_4)\}
  \label{eq:branch-free-energy-asymptotics}
\end{equation}
uniformly over the finite branch set, then
\begin{equation}
  \lim_{\Lambda\to\infty}\lim_{V_4\to\infty}
  p_{\Lambda,V_4,\beta_*}(\theta)=1,
  \qquad
  \lim_{\Lambda\to\infty}\lim_{V_4\to\infty}
  p_{\Lambda,V_4,\beta}(\theta)=0
  \quad(\beta\neq\beta_*).
  \label{eq:theta-branch-selection}
\end{equation}
For \(\beta\neq\beta_*\),
\[
  \frac{Z_{\Lambda,V_4,\beta}(\theta)}
       {Z_{\Lambda,V_4,\beta_*}(\theta)}
  =
  \exp\{-V_4(E_\beta(\theta)-E_{\beta_*}(\theta))+o(V_4)\}.
\]
The exponent tends to \(-\infty\) because the energy-density gap is positive.
Since the branch set is finite, the sum of all nonminimal ratios tends to
zero.  Dividing the branch weights by
\(Z_{\Lambda,V_4,\beta_*}\) gives
\eqref{eq:theta-branch-selection}.
\end{remark}

\begin{remark}[Cluster test for mixtures of theta branches]
\label{rem:theta-branch-mixture-cluster-test}
Let \(\{\omega_\beta\}_{\beta\in\mathcal B}\) be translation-invariant pure
states for gauge-invariant local observables, and assume each branch clusters:
for local observables \(\mathcal O,\mathcal P\),
\begin{equation}
  \lim_{|a|\to\infty}
  \omega_\beta(\mathcal O(a)\mathcal P(0))
  =
  \omega_\beta(\mathcal O)\omega_\beta(\mathcal P).
  \label{eq:branch-clustering-assumption}
\end{equation}
For a convex combination
\begin{equation}
  \omega=\sum_{\beta\in\mathcal B}p_\beta\omega_\beta,
  \qquad
  p_\beta\ge0,
  \qquad
  \sum_\beta p_\beta=1,
  \label{eq:theta-branch-mixture-state}
\end{equation}
one has
\begin{equation}
  \lim_{|a|\to\infty}
  \left[
    \omega(\mathcal O(a)\mathcal P(0))
    -
    \omega(\mathcal O)\omega(\mathcal P)
  \right]
  =
  \operatorname{Cov}_{p}
  \bigl(\omega_\beta(\mathcal O),\omega_\beta(\mathcal P)\bigr).
  \label{eq:theta-branch-cluster-covariance}
\end{equation}
Thus the mixture clusters for all local observables only if the branch labels
cannot be distinguished by local one-point functions on the support of
\(p_\beta\), or if the mixture is supported on a single branch.
Using the branch clustering assumption,
\[
  \lim_{|a|\to\infty}
  \omega(\mathcal O(a)\mathcal P(0))
  =
  \sum_\beta p_\beta
  \omega_\beta(\mathcal O)\omega_\beta(\mathcal P).
  \]
The product of the one-point functions in the mixed state is
\[
  \omega(\mathcal O)\omega(\mathcal P)
  =
  \left(\sum_\beta p_\beta\omega_\beta(\mathcal O)\right)
  \left(\sum_\gamma p_\gamma\omega_\gamma(\mathcal P)\right).
  \]
Their difference is precisely the covariance of the two branch-dependent
one-point functions with respect to the probability distribution \(p\).
\end{remark}

\begin{remark}[Theta vacua as theorem, hypothesis, and convention]
\label{rem:theta-vacua-status}
At finite regulator the sector decomposition, cumulant identity, and branch
mixture algebra above are exact.  A continuum statement that theta labels
superselection sectors requires additional input: the existence of the
continuum gauge theory, the infinite-volume limits of the local states, the
cluster property inside a selected branch, and the action of large gauge
transformations or theta-shift automorphisms on the physical representation.
The monograph therefore uses ``theta vacuum'' only after these data have been
specified.  Otherwise it speaks of a theta parameter in the regulated action
or of a theta-branch hypothesis.
\end{remark}

\section{The BPST Instanton and the Semiclassical Vertex}
\label{sec:bpst-instanton-thooft-vertex}

We now construct the basic finite-action representative of the sector
\(Q=1\).  This section is Euclidean: coordinates are
\(x^\mu\), \(\mu=1,\ldots,4\), the metric is \(\delta_{\mu\nu}\), and
\(\epsilon_{1234}=+1\).  The chapter's gauge-field convention remains
Hermitian, so
\[
  F_{\mu\nu}
  =
  \partial_\mu A_\nu-\partial_\nu A_\mu-\ii[A_\mu,A_\nu].
\]
For explicit \(SU(2)\) formulas it is economical to use the auxiliary basis
\[
  T_a=\frac{\sigma_a}{2},
  \qquad
  [T_a,T_b]=\ii\epsilon^c{}_{ab}T_c,
  \qquad
  \operatorname{tr}(T_aT_b)=\frac12\delta_{ab}.
\]
This is only a coordinate choice for the displayed solution.  The monograph's
default trace-delta basis is \(t_a=\sqrt2\,T_a\).  If
\[
  \mathcal L_{\rm here}
  =
  -\frac1{4g_{\rm YM}^2}
  \operatorname{tr}_{\delta}(F_{\mu\nu}F^{\mu\nu})
\]
is rewritten in the \(T_a\) basis, then the corresponding half-trace coupling
is \(g_{\rm ht}=\sqrt2\,g_{\rm YM}\).  Thus the familiar instanton action
\(8\pi^2/g_{\rm ht}^2\) is the same number as
\(4\pi^2/g_{\rm YM}^2\) in the active trace-delta coupling coordinate.

Define the self-dual 't Hooft symbols by
\[
  \eta^a_{ij}=\epsilon_{aij},
  \qquad
  \eta^a_{i4}=\delta_{ai},
  \qquad
  \eta^a_{4i}=-\delta_{ai},
  \qquad i,j,a=1,2,3,
\]
so that
\[
  \eta^a_{\mu\nu}
  =
  \frac12\epsilon_{\mu\nu\rho\sigma}\eta^a_{\rho\sigma},
  \qquad
  \sum_{a,\mu,\nu}\eta^a_{\mu\nu}\eta^a_{\mu\nu}=12 .
\]
The contraction identity used below is
\begin{equation}
  \epsilon^a{}_{bc}\eta^b_{\mu\rho}\eta^c_{\nu\sigma}
  y^\rho y^\sigma
  =
  r^2\eta^a_{\mu\nu}
  +y_\mu\eta^a_{\nu\lambda}y^\lambda
  -y_\nu\eta^a_{\mu\lambda}y^\lambda .
  \label{eq:eta-symbol-quadratic-identity}
\end{equation}
It follows by checking the three possible positions of the index \(4\) in the
defining equations for \(\eta\), and then using antisymmetry in
\(\mu,\nu\).  This identity is the finite-dimensional algebraic input that
turns the apparently inhomogeneous derivative and commutator terms in the
curvature into a multiple of \(\eta^a_{\mu\nu}\).
Let \(a^\mu\in\mathbb R^4\) and \(\rho>0\).  Put \(y=x-a\) and
\(r^2=y_\mu y_\mu\).  The BPST one-instanton in regular gauge is
\begin{equation}
  A_\mu(x)
  =
  \frac{2\,\eta^a_{\mu\nu}y^\nu}{r^2+\rho^2}\,T_a .
  \label{eq:bpst-regular-gauge-potential}
\end{equation}
Write \(f=r^2+\rho^2\) and \(A_\mu=A_\mu^aT_a\).  Since
\[
  \partial_\mu A_\nu^a-\partial_\nu A_\mu^a
  =
  -\frac4f\,\eta^a_{\mu\nu}
  -
  \frac4{f^2}
  \left(
    y_\mu\eta^a_{\nu\lambda}y^\lambda
    -
    y_\nu\eta^a_{\mu\lambda}y^\lambda
  \right)
\]
and
\[
  \epsilon^a{}_{bc}A_\mu^bA_\nu^c
  =
  \frac4{f^2}
  \left(
    r^2\eta^a_{\mu\nu}
    +
    y_\mu\eta^a_{\nu\lambda}y^\lambda
    -
    y_\nu\eta^a_{\mu\lambda}y^\lambda
  \right),
\]
the curvature convention
\(F^a_{\mu\nu}
=\partial_\mu A_\nu^a-\partial_\nu A_\mu^a+\epsilon^a{}_{bc}A_\mu^bA_\nu^c\)
gives
\begin{equation}
  F_{\mu\nu}(x)
  =
  -\frac{4\rho^2}{(r^2+\rho^2)^2}\,
  \eta^a_{\mu\nu}T_a .
  \label{eq:bpst-regular-gauge-curvature}
\end{equation}
Hence \(F_{\mu\nu}=\widetilde F_{\mu\nu}\), where
\(\widetilde F_{\mu\nu}=\frac12\epsilon_{\mu\nu\rho\sigma}F_{\rho\sigma}\).
The connection is smooth at \(r=0\).  At infinity it is gauge-equivalent to a
pure gauge of winding one.  Equivalently, applying the gauge transformation
\[
  g(y)=\frac{y_4\mathbf 1+\ii y_i\sigma_i}{r},
  \qquad g:S^3_\infty\to SU(2),
\]
produces the singular-gauge representative
\[
  A_\mu^{\rm sing}(x)
  =
  \frac{2\rho^2\,\bar\eta^a_{\mu\nu}y^\nu}
       {r^2(r^2+\rho^2)}\,T_a ,
\]
where \(\bar\eta\) is the anti-self-dual 't Hooft symbol obtained from
\(\eta\) by reversing the signs of the components with one index equal to
\(4\).  This representative decays as \(r^{-3}\) at infinity and has a
gauge-coordinate
singularity at \(y=0\).  The two representatives describe the same smooth
connection on the one-instanton bundle over \(S^4\).

The action saturation is an algebraic consequence of self-duality.  In the
half-trace component convention,
\[
  S_E^{\rm ht}
  =
  \frac1{4g_{\rm ht}^2}
  \int_{\mathbb R^4}\dd^4x\,
  F^a_{\mu\nu}F^a_{\mu\nu}.
\]
Completing the square gives, for any finite-action field in the sector of
charge \(k\),
\begin{equation}
  S_E^{\rm ht}
  =
  \frac1{8g_{\rm ht}^2}
  \int (F^a_{\mu\nu}\mp\widetilde F^a_{\mu\nu})^2\,\dd^4x
  \pm
  \frac1{4g_{\rm ht}^2}
  \int F^a_{\mu\nu}\widetilde F^a_{\mu\nu}\,\dd^4x
  \ge
  \frac{8\pi^2}{g_{\rm ht}^2}|k|.
  \label{eq:instanton-action-bound-half-trace}
\end{equation}
Equality holds precisely for self-dual fields when \(k>0\) and anti-self-dual
fields when \(k<0\).  For \eqref{eq:bpst-regular-gauge-curvature},
\[
  F^a_{\mu\nu}F^a_{\mu\nu}
  =
  \frac{192\rho^4}{(r^2+\rho^2)^4},
  \qquad
  \int_{\mathbb R^4}
  \frac{\rho^4\,\dd^4x}{(r^2+\rho^2)^4}
  =
  \frac{\pi^2}{6},
\]
where the last integral is
\[
  2\pi^2\int_0^\infty
  \frac{\rho^4r^3\,\dd r}{(r^2+\rho^2)^4}
  =
  \pi^2\int_0^\infty
  \frac{\rho^4u\,\dd u}{(u+\rho^2)^4}
  =
  \frac{\pi^2}{6}.
\]
Therefore
\[
  \int_{\mathbb R^4}F^a_{\mu\nu}F^a_{\mu\nu}\,\dd^4x=32\pi^2,
  \qquad
  S_E^{\rm ht}=\frac{8\pi^2}{g_{\rm ht}^2}.
\]
Because the field is self-dual,
\[
  Q[A]
  =
  \frac1{32\pi^2}
  \int F^a_{\mu\nu}\widetilde F^a_{\mu\nu}\,\dd^4x
  =
  1 .
\]
In the trace-delta coupling coordinate used elsewhere in this monograph this
same saddle contributes the exponential
\begin{equation}
  \exp\!\left(-\frac{4\pi^2}{g_{\rm YM}^2}+\ii\theta\right)
  =
  \exp\!\left(-\frac{8\pi^2}{g_{\rm ht}^2}+\ii\theta\right).
  \label{eq:bpst-instanton-exponential-conventions}
\end{equation}

\begin{lemma}[BPST one-instanton data]
\label{lem:bpst-one-instanton-data}
With the self-dual symbols and \(SU(2)\) half-trace generators
\(T_a=\sigma_a/2\) used in this section, the connection
\[
  A_\mu(x)
  =
  \frac{2\,\eta^a_{\mu\nu}(x-a)^\nu}{(x-a)^2+\rho^2}\,T_a,
  \qquad
  \rho>0,
\]
has curvature
\[
  F_{\mu\nu}(x)
  =
  -\frac{4\rho^2}{((x-a)^2+\rho^2)^2}\,
  \eta^a_{\mu\nu}T_a .
\]
It is self-dual, has topological charge \(Q=1\), and has Euclidean action
\[
  S_E^{\rm ht}=\frac{8\pi^2}{g_{\rm ht}^2}
  =
  \frac{4\pi^2}{g_{\rm YM}^2}
\]
when rewritten in the trace-delta coupling coordinate of the Yang--Mills
chapters.
\end{lemma}

\begin{proof}
Differentiating the displayed connection gives the derivative term written
above, and the commutator term is
\(\epsilon^a{}_{bc}A_\mu^bA_\nu^c\).  The contraction identity
\eqref{eq:eta-symbol-quadratic-identity} cancels the inhomogeneous
\(y_\mu\eta_{\nu\lambda}^a y^\lambda
-y_\nu\eta_{\mu\lambda}^a y^\lambda\) terms and leaves
\eqref{eq:bpst-regular-gauge-curvature}.  Since
\(\eta^a_{\mu\nu}
=\frac12\epsilon_{\mu\nu\rho\sigma}\eta^a_{\rho\sigma}\), this curvature is
self-dual.

The square-completion identity
\eqref{eq:instanton-action-bound-half-trace} gives the lower bound
\(8\pi^2|Q|/g_{\rm ht}^2\), saturated by a self-dual field with \(Q>0\).
For the displayed curvature,
\[
  F^a_{\mu\nu}F^a_{\mu\nu}
  =
  \frac{192\rho^4}{(r^2+\rho^2)^4},
  \qquad
  \int_{\mathbb R^4}
  \frac{\rho^4\,\dd^4x}{(r^2+\rho^2)^4}
  =
  \frac{\pi^2}{6},
\]
so \(\int F^a_{\mu\nu}F^a_{\mu\nu}=32\pi^2\).  Self-duality then gives
\[
  Q=\frac1{32\pi^2}\int F^a_{\mu\nu}\widetilde F^a_{\mu\nu}\,\dd^4x=1
\]
and \(S_E^{\rm ht}=8\pi^2/g_{\rm ht}^2\).  Finally,
\(\operatorname{tr}_\delta(t_at_b)=\delta_{ab}\) with
\(t_a=\sqrt2\,T_a\) implies \(g_{\rm ht}=\sqrt2\,g_{\rm YM}\), hence the
trace-delta action is \(4\pi^2/g_{\rm YM}^2\).
\end{proof}

Figure~\ref{fig:bpst-instanton-data} plots the radial content of the same
normalization check.  In the dimensionless variable \(u=r/\rho\), the
normalized action-density profile and the cumulative topological charge are
\[
  \frac{F^a_{\mu\nu}F^a_{\mu\nu}(r)}
       {F^a_{\mu\nu}F^a_{\mu\nu}(0)}
  =
  \frac1{(1+u^2)^4},
  \qquad
  \frac1{32\pi^2}\int_{|x-a|\le R}
  F^a_{\mu\nu}F^a_{\mu\nu}\,\dd^4x
  =
  1-\frac3{(1+u^2)^2}+\frac2{(1+u^2)^3}.
\]
The second formula is the radial integral in the proof of
Lemma~\ref{lem:bpst-one-instanton-data} before the limit
\(R/\rho\to\infty\).

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1.2cm,y=3.2cm,every node/.style={font=\scriptsize}]
    \draw[->] (-0.08,0) -- (5.45,0) node[below right] {\(u=r/\rho\)};
    \draw[->] (0,-0.05) -- (0,1.15) node[left] {normalized value};
    \foreach \x in {1,2,3,4}
      \draw (\x,0.015) -- (\x,-0.015) node[below] {\(\x\)};
    \foreach \y/\lab in {0.5/{1/2},1/{1}}
      \draw (0.018,\y) -- (-0.018,\y) node[left] {\(\lab\)};
    \draw[densely dotted] (1,0) -- (1,1.04);
    \node[anchor=north west] at (1.03,-0.06) {\(u=1\)};

    \draw[blue!70!black,thick] plot[domain=0:5,samples=120]
      ({\x},{1/(1+\x*\x)/(1+\x*\x)/(1+\x*\x)/(1+\x*\x)});
    \draw[red!70!black,thick] plot[domain=0:5,samples=120]
      ({\x},{1 - 3/(1+\x*\x)/(1+\x*\x)
        + 2/(1+\x*\x)/(1+\x*\x)/(1+\x*\x)});
    \draw[red!70!black,densely dotted] (0,1) -- (5,1);

    \node[blue!70!black,align=left,anchor=west] at (1.45,0.35)
      {\(\displaystyle (1+u^2)^{-4}\)\\field-strength density};
    \node[red!70!black,align=left,anchor=west] at (2.10,0.84)
      {\(\displaystyle 1-\frac3{(1+u^2)^2}
        +\frac2{(1+u^2)^3}\)\\cumulative \(Q(R)\)};
    \node[align=center] at (2.65,-0.27)
      {the instanton core has size \(\rho\), while \(Q(R)\to1\)};
  \end{tikzpicture}
  }
  \caption{Radial BPST instanton profile in units of its size \(\rho\).  The
  blue curve is the normalized density \(F^a_{\mu\nu}F^a_{\mu\nu}(r)/
  F^a_{\mu\nu}F^a_{\mu\nu}(0)\); the red curve is the cumulative
  topological charge inside radius \(R=\rho u\).  The plot makes the
  localized core and the normalization \(Q=1\) visible rather than merely
  repeating the algebraic formulas.}
  \label{fig:bpst-instanton-data}
\end{figure}

The bosonic zero modes are tangent vectors to the moduli of solutions modulo
gauge transformations.  For \(SU(2)\), \(k=1\), the smooth moduli are
\[
  a^\mu\in\mathbb R^4,\qquad
  \rho\in\mathbb R_{>0},\qquad
  U\in SU(2)/\mathbb Z_2 .
\]
Thus there are
\[
  4+1+3=8
\]
bosonic zero modes: four translations, one dilation, and three global gauge
orientations.  For an \(SU(N_c)\) instanton obtained by embedding the above
\(SU(2)\) solution, the orientation moduli are enlarged, and the
one-instanton bosonic moduli space has dimension \(4N_c\) before quotienting
possible global identifications.  The reason for the number \(4N_c\) is
local and elementary.  The center and size contribute \(4+1\) parameters.  The
embedded \(SU(2)\) orientation is the orbit of the embedding under
\(SU(N_c)\), whose stabilizer is \(SU(N_c-2)\times U(1)\).  Hence the
orientation dimension is
\[
  (N_c^2-1)-\bigl((N_c-2)^2-1\bigr)-1=4N_c-5 ,
\]
and adding center and size gives \(4N_c\).  The precise global quotient by
the center and by stabilizers affects the coordinate range, not this local
dimension count.

\begin{definition}[Framed ADHM quotient]
\label{def:framed-adhm-quotient-datum}
Let \(W\simeq\mathbb C^{N_c}\) be the framing vector space and
\(K\simeq\mathbb C^k\) the instanton vector space.  A framed \(U(N_c)\)
ADHM quadruple of charge \(k\) is
\[
  (B_1,B_2,I,J)
  \in
  \operatorname{End}(K)^2
  \oplus\operatorname{Hom}(W,K)
  \oplus\operatorname{Hom}(K,W),
\]
with moment maps
\begin{align}
  \mu_{\mathbb C}
  &:=
  [B_1,B_2]+IJ
  \in\operatorname{End}(K),
  \nonumber\\
  \mu_{\mathbb R}
  &:=
  [B_1,B_1^\dagger]+[B_2,B_2^\dagger]
  +II^\dagger-J^\dagger J
  \in\operatorname{Herm}(K).
  \label{eq:general-adhm-moment-maps}
\end{align}
The unitary group \(U(K)\) acts by
\begin{equation}
  (B_1,B_2,I,J)
  \longmapsto
  (gB_1g^{-1},gB_2g^{-1},gI,Jg^{-1}).
  \label{eq:general-adhm-unitary-action}
\end{equation}
The smooth framed ADHM quotient is the stable part of
\begin{equation}
  \mathcal M_{k,N_c}^{\rm ADHM}
  :=
  \{\,\mu_{\mathbb C}=0,\ \mu_{\mathbb R}=0\,\}/U(K).
  \label{eq:general-adhm-quotient}
\end{equation}
Here stability means that there is no proper subspace
\(S\subset K\) which contains \(I(W)\) and is preserved by both \(B_1\) and
\(B_2\).  On this stable locus the framed gauge data have no residual
endomorphism of \(K\) invisible to the framing.
\end{definition}

The ADHM construction identifies the smooth quotient
\eqref{eq:general-adhm-quotient} with the framed smooth instanton moduli
space on \(\mathbb R^4\), after matching the chosen orientation convention
for self-duality versus anti-self-duality.  The full transform also builds
the connection from the finite-dimensional linear algebra.  The calculation
needed for the semiclassical measure at this stage is the local quotient
dimension and the boundary behavior of the quotient, which follow directly
from Definition~\ref{def:framed-adhm-quotient-datum}.

\begin{proposition}[ADHM quotient dimension at smooth stable points]
\label{prop:adhm-quotient-dimension-count}
At a point of \(\mu_{\mathbb C}^{-1}(0)\cap\mu_{\mathbb R}^{-1}(0)\) where
the stable \(U(K)\) action is free and the moment map is transverse to zero,
the framed ADHM quotient
\(\mathcal M_{k,N_c}^{\rm ADHM}\) has local real dimension
\[
  \dim_\mathbb R\mathcal M_{k,N_c}^{\rm ADHM}=4kN_c .
\]
The trace parts of \(B_1\) and \(B_2\) give the four center coordinates of
the instanton configuration.  The centered quotient therefore has local real
dimension \(4kN_c-4\).
\end{proposition}

\begin{proof}
The real dimension of the unconstrained variables is
\[
  2\dim_\mathbb C\bigl(\operatorname{End}(K)^2\bigr)
  +
  2\dim_\mathbb C\operatorname{Hom}(W,K)
  +
  2\dim_\mathbb C\operatorname{Hom}(K,W)
  =
  4k^2+4kN_c .
\]
The complex equation \(\mu_{\mathbb C}=0\) imposes \(k^2\) complex
conditions, hence \(2k^2\) real conditions.  The real moment-map equation
\(\mu_{\mathbb R}=0\) imposes \(k^2\) real conditions because
\(\mathfrak u(K)\) has real dimension \(k^2\).  Transversality makes these
conditions independent at the point under consideration.

Finally one divides by \(U(K)\), whose real dimension is \(k^2\).  Freeness
on the stable locus means that this quotient subtracts exactly \(k^2\)
dimensions locally.  Thus
\[
  (4k^2+4kN_c)-2k^2-k^2-k^2=4kN_c .
\]
The equations contain only commutators of \(B_1,B_2\) and the products
\(IJ,II^\dagger,J^\dagger J\).  Shifting
\[
  B_\alpha\longmapsto B_\alpha+c_\alpha\mathbf 1_K,
  \qquad \alpha=1,2,
\]
leaves the moment maps unchanged.  The two complex numbers
\(\operatorname{tr}B_1/k\) and \(\operatorname{tr}B_2/k\) are therefore the
four real translation coordinates in the complex structure used to write
\(\mathbb R^4\simeq\mathbb C^2\).  Removing them gives centered dimension
\(4kN_c-4\).
\end{proof}

The stability condition is also where the small-instanton boundary first
appears algebraically.  If a stable ADHM tuple degenerates so that a nonzero
subspace of \(K\) decouples from the framing, the quotient leaves the smooth
instanton stratum and records instanton charge concentrated at points of
\(\mathbb R^4\).  The Uhlenbeck boundary has strata of the form
\[
  \mathcal M_{k-\ell,N_c}^{\rm smooth}
  \times
  \operatorname{Sym}^{\ell}(\mathbb R^4),
  \qquad 1\le \ell\le k,
\]
where \(\ell\) units of charge have collapsed to pointlike ideal instantons.
For a path integral this compactification carries analytic content: whether the
semiclassical integral is finite depends on the determinant, operator
insertions, and regulator behavior near these boundary strata.

\begin{definition}[ADHM quotient density]
\label{def:adhm-quotient-density}
Equip the affine ADHM space
\[
  \mathbb A_{k,N_c}
  =
  \operatorname{End}(K)^2
  \oplus\operatorname{Hom}(W,K)
  \oplus\operatorname{Hom}(K,W)
\]
with the flat Hermitian metric
\[
  \|(\delta B_1,\delta B_2,\delta I,\delta J)\|^2
  =
  \operatorname{tr}_K(\delta B_1\delta B_1^\dagger)
  +\operatorname{tr}_K(\delta B_2\delta B_2^\dagger)
  +\operatorname{tr}_K(\delta I\delta I^\dagger)
  +\operatorname{tr}_W(\delta J^\dagger\delta J).
\]
On the free, stable, regular locus
\[
  Z_{k,N_c}^{\rm reg}
  :=
  \{\mu_{\mathbb C}=0,\ \mu_{\mathbb R}=0\}_{\rm stable,reg}
  \subset \mathbb A_{k,N_c},
\]
let \(V_X\subset T_X Z_{k,N_c}^{\rm reg}\) be the tangent space to the
\(U(K)\)-orbit and let
\[
  H_X:=V_X^\perp\cap T_XZ_{k,N_c}^{\rm reg}.
\]
The differential of the quotient map identifies \(H_X\) with the tangent
space of the smooth quotient at \([X]\).  The \emph{ADHM quotient density}
\(\dd\nu_{k,N_c}^{\rm ADHM}\) is the Riemannian density induced by this
quotient metric.
\end{definition}

This density is a finite-dimensional object.  It is not yet the complete
semiclassical Yang--Mills weight: the latter also contains the nonzero-mode
determinants, coupling-coordinate convention, operator renormalization, and
the treatment of the boundary strata.  The quotient density is the part of
the measure fixed by the classical zero-mode geometry.

\begin{proposition}[Coarea form of the ADHM quotient density]
\label{prop:adhm-quotient-density-coarea}
Fix an \(\operatorname{Ad}\)-invariant inner product on
\(\mathfrak u(K)\) and let \(\dd\operatorname{vol}_{U(K)}\) be the
associated left-invariant Riemannian density on \(U(K)\).  Choose an
orthonormal basis \(\{\xi_A\}_{A=1}^{k^2}\) of \(\mathfrak u(K)\), and let
\(\xi_A^\#(X)\) denote the corresponding infinitesimal orbit vector at
\(X\in Z_{k,N_c}^{\rm reg}\).  Define the orbit Gram matrix
\[
  M_{AB}(X):=\langle \xi_A^\#(X),\xi_B^\#(X)\rangle_{\mathbb A}.
\]
If \(F\) is a compactly supported smooth \(U(K)\)-invariant function on
\(Z_{k,N_c}^{\rm reg}\), then
\begin{equation}
  \int_{\mathcal M_{k,N_c}^{\rm ADHM}}
  F([X])\,\dd\nu_{k,N_c}^{\rm ADHM}([X])
  =
  \frac1{\operatorname{Vol}U(K)}
  \int_{Z_{k,N_c}^{\rm reg}}
  \frac{F(X)}{\sqrt{\det M(X)}}\,\dd\operatorname{vol}_{Z}(X).
  \label{eq:adhm-quotient-density-levelset-coarea}
\end{equation}
Equivalently, if \(J_\mu(X)\) is the coarea Jacobian
\[
  J_\mu(X)
  :=
  \det\!\bigl(D\mu_XD\mu_X^\ast\bigr)^{1/2},
  \qquad
  \mu=(\operatorname{Re}\mu_{\mathbb C},
       \operatorname{Im}\mu_{\mathbb C},\mu_{\mathbb R}),
\]
then the same density may be represented distributionally on the ambient
affine space by
\begin{equation}
  \int_{\mathcal M_{k,N_c}^{\rm ADHM}}
  F\,\dd\nu_{k,N_c}^{\rm ADHM}
  =
  \frac1{\operatorname{Vol}U(K)}
  \int_{\mathbb A_{k,N_c}}
  F(X)\,
  \delta(\mu(X))\,
  \frac{J_\mu(X)}{\sqrt{\det M(X)}}\,
  \dd X ,
  \label{eq:adhm-quotient-density-ambient-coarea}
\end{equation}
on any compact set where the regularity and freeness hypotheses hold.  The
delta function in \eqref{eq:adhm-quotient-density-ambient-coarea} is
normalized with respect to the Euclidean density on the real target of
\(\mu\) induced by the same trace pairings.
\end{proposition}

\begin{proof}
The restricted flat metric on \(Z_{k,N_c}^{\rm reg}\) gives the orthogonal
splitting
\[
  T_XZ_{k,N_c}^{\rm reg}=H_X\oplus V_X.
\]
In local product coordinates \((y,\theta)\), where \(y\) is a quotient
coordinate and \(\theta^A\) are group coordinates near the identity, the
metric is block diagonal at the chosen slice if \(H_X\) is used as the
horizontal distribution.  The horizontal block is the quotient metric, while
the vertical block is precisely \(M_{AB}(X)\).  Therefore the Riemannian
density on the level set factorizes as
\[
  \dd\operatorname{vol}_{Z}
  =
  \sqrt{\det M(X)}\,
  \dd\operatorname{vol}_{U(K)}\,
  \dd\nu_{k,N_c}^{\rm ADHM}.
\]
Because the flat metric and the moment-map equations are \(U(K)\)-invariant,
\(\det M(X)\) is constant along each orbit.  Integrating an invariant
function over the group coordinate gives
\eqref{eq:adhm-quotient-density-levelset-coarea}.

For the ambient-space expression, apply the ordinary finite-dimensional
coarea formula to the submersion \(\mu\):
\[
  \int_{\mathbb A_{k,N_c}}
  h(X)\,\delta(\mu(X))\,J_\mu(X)\,\dd X
  =
  \int_{\mu^{-1}(0)} h(X)\,\dd\operatorname{vol}_{Z}(X)
\]
for compactly supported \(h\) in the regular locus.  Taking
\(h=F/\sqrt{\det M}\) gives
\eqref{eq:adhm-quotient-density-ambient-coarea}.
\end{proof}

\begin{remark}[Classical zero-mode density versus determinant weight]
\label{rem:adhm-density-versus-determinants}
Equations
\eqref{eq:adhm-quotient-density-levelset-coarea} and
\eqref{eq:adhm-quotient-density-ambient-coarea} define the classical
zero-mode density on the smooth ADHM quotient.  They do not decide whether
the instanton contribution to a continuum QFT observable is finite near a
small-instanton stratum.  That question depends on the nonzero-mode
determinant, the operator insertions, and the renormalization scheme.  This
is why the Uhlenbeck boundary is part of the measure problem rather than an
afterthought about topology.
\end{remark}

\begin{proposition}[Charge-one ADHM chart and the instanton orientation orbit]
\label{prop:k-one-adhm-chart-orientation-orbit}
Let \(W\simeq\mathbb C^{N_c}\) with its Hermitian inner product and let
\(K\simeq\mathbb C\).  The centered charge-one framed \(U(N_c)\) ADHM data
are a pair
\[
  I\in\operatorname{Hom}(W,K),
  \qquad
  J\in\operatorname{Hom}(K,W),
\]
obeying
\begin{equation}
  IJ=0,
  \qquad
  II^\dagger-J^\dagger J=0,
  \label{eq:k-one-adhm-moment-equations}
\end{equation}
with \((I,J)\ne(0,0)\), modulo the \(U(1)\) action
\[
  (I,J)\longmapsto(e^{\ii\alpha}I,e^{-\ii\alpha}J).
\]
After identifying \(I\) with a vector \(u\in W\) by the Hermitian metric and
writing \(J\) as a vector \(v\in W\), the equations are
\begin{equation}
  \langle u,v\rangle=0,
  \qquad
  \|u\|=\|v\|=\rho>0 .
  \label{eq:k-one-adhm-orthonormal-pair}
\end{equation}
For fixed \(\rho\), the quotient is the homogeneous space
\begin{equation}
  V_2(\mathbb C^{N_c})/U(1)
  \simeq
  U(N_c)/(U(N_c-2)\times U(1)),
  \label{eq:k-one-adhm-orientation-orbit}
\end{equation}
where \(V_2(\mathbb C^{N_c})\) is the Stiefel manifold of ordered
orthonormal two-frames, \(U(0)\) is interpreted as the trivial group when
\(N_c=2\), and the \(U(1)\) in the quotient acts by the common phase of the
two frame vectors.  Equivalently, for \(N_c\ge3\), after
restricting to the special unitary framing, the connected orbit has the same
local form
\[
  SU(N_c)/(SU(N_c-2)\times U(1))
\]
up to a finite central quotient.  For \(N_c=2\) this specialization is the
usual \(SU(2)/\mathbb Z_2\) orientation orbit.  In every case the real
dimension is \(4N_c-5\).  Adding the size \(\rho\) and the four center
coordinates gives the local real dimension \(4N_c\) of the charge-one
instanton moduli space.
\end{proposition}

\begin{proof}
For \(K\simeq\mathbb C\), the ADHM endomorphisms \(B_1,B_2\in
\operatorname{End}(K)\) are complex numbers.  They commute and therefore
contribute no centered ADHM equation; their real and imaginary parts are the
four center coordinates of the instanton.  The remaining ADHM equations are
exactly \eqref{eq:k-one-adhm-moment-equations}: \(IJ=0\) is the complex
moment-map equation and \(II^\dagger-J^\dagger J=0\) is the real moment-map
equation for the \(U(1)\) acting on \(K\).

Use the Hermitian metric to identify the row \(I\) with a vector \(u\in W\)
by \(I(w)=\langle u,w\rangle\), and identify \(J\) with \(v=J(1)\).  Then
\(IJ=I(v)=\langle u,v\rangle\), while
\[
  II^\dagger=\|u\|^2\operatorname{id}_{K},
  \qquad
  J^\dagger J=\|v\|^2\operatorname{id}_{K}.
\]
Thus \eqref{eq:k-one-adhm-moment-equations} is equivalent to
\eqref{eq:k-one-adhm-orthonormal-pair}.  Since \((I,J)\ne(0,0)\), the common
norm is a positive number \(\rho\).  Dividing by \(\rho\) gives an ordered
orthonormal two-frame \((u/\rho,v/\rho)\).

The \(U(1)\) ADHM quotient sends
\((u,v)\mapsto(e^{-\ii\alpha}u,e^{-\ii\alpha}v)\), namely the common phase of
the ordered frame.  The unitary group \(U(N_c)\) acts transitively on ordered
orthonormal two-frames.  The stabilizer of the common-phase equivalence class
of the reference frame \((e_1,e_2)\) consists of an arbitrary unitary
transformation on the orthogonal complement together with a common phase on
the two displayed vectors, hence is \(U(N_c-2)\times U(1)\).  This proves
\eqref{eq:k-one-adhm-orientation-orbit}.  Its dimension is
\[
  \dim_\mathbb R U(N_c)
  -
  \dim_\mathbb R U(N_c-2)
  -
  \dim_\mathbb R U(1)
  =
  N_c^2-(N_c-2)^2-1
  =
  4N_c-5 .
\]
The size contributes one more dimension and the center contributes four.
\end{proof}

\paragraph{Cone structure and the small-instanton boundary.}
On the positive-size charge-one stratum, the centered ADHM quotient is the
metric cone over the orientation orbit
\[
  \mathcal O_{N_c}
  =
  U(N_c)/(U(N_c-2)\times U(1)).
\]
With the quotient metric induced from the flat metric
\(\|dI\|^2+\|dJ\|^2\), its volume density has the local radial form
\begin{equation}
  \rho^{4N_c-5}\,\dd\rho\,\dd\Omega_{N_c},
  \label{eq:k-one-adhm-classical-cone-density}
\end{equation}
for an invariant density \(\dd\Omega_{N_c}\) on \(\mathcal O_{N_c}\).  The
endpoint \(\rho=0\) is the small-instanton boundary: the two-frame collapses
to \((I,J)=(0,0)\), the stabilizer jumps, and the positive-size quotient no
longer supplies a smooth manifold chart.
The equations \eqref{eq:k-one-adhm-orthonormal-pair} are homogeneous in
\((u,v)\).  Hence every positive-size solution is uniquely written as
\(\rho(\hat u,\hat v)\), with \((\hat u,\hat v)\in V_2(\mathbb C^{N_c})\)
and \(\rho>0\), modulo the same common phase.  The flat metric on
\((I,J)\) scales as
\[
  \|dI\|^2+\|dJ\|^2
  =
  c\,\dd\rho^2+\rho^2\,\dd s^2_{\mathcal O_{N_c}}
\]
for a positive constant \(c\) depending only on the normalization of
\(\rho\).  A metric cone over a \(d\)-dimensional base has radial volume
density proportional to \(\rho^d\dd\rho\) times the base density.  Here
\(d=\dim_\mathbb R\mathcal O_{N_c}=4N_c-5\), giving
\eqref{eq:k-one-adhm-classical-cone-density}.  At \(\rho=0\) the frame
vectors vanish and no two-frame remains; this is precisely the
finite-dimensional ADHM model of the small-instanton degeneration.

The regulated collective-coordinate measure is the one constructed in
Theorem~\ref{thm:collective-coordinate-morse-bott-measure}.  In a finite
gauge-fixed regulator its local bosonic factor has the form
\begin{equation}
  \dd\mu_{k=1,\Lambda}^{\rm bos}
  =
  \frac{\Delta_{\rm FP,\Lambda}(A_z)}
       {\operatorname{Vol}(\Gamma_z)}
  \frac{\sqrt{\det G^{(\Lambda)}_{ab}(z)}\,
        \dd^4a\,\dd\rho\,
        \dd\Omega_{SU(N_c)/(SU(N_c-2)\times U(1))}}
       {\det{}'_\Lambda(L_{A_z})^{1/2}},
  \label{eq:one-instanton-bosonic-collective-measure}
\end{equation}
multiplied by the exponential
\eqref{eq:bpst-instanton-exponential-conventions}.  Here \(z\) denotes the
center, size, and orientation coordinates; \(G^{(\Lambda)}_{ab}\) is the
zero-mode Gram matrix induced by the regulated gauge-field inner product;
\(L_{A_z}\) is the gauge-fixed quadratic operator on nonzero bosonic
fluctuations; and \(\Gamma_z\) is the residual stabilizer left by the chosen
slice.  Equation
\eqref{eq:one-instanton-bosonic-collective-measure} is not yet a continuum
formula.  Taking \(\Lambda\to\infty\) requires combining the determinant, the
coupling coordinate, and the normalization of the local operator insertions
in a specified renormalization scheme.  What is scheme independent at this
stage is the topological sector, the zero-mode count, the finite-regulator
construction of the measure, and the exponential action.

\paragraph{Finite-regulator nonzero-mode determinants.}
The symbol ``determinant constant'' hides several different objects unless
the regulator is kept explicit.  At fixed \(\Lambda\), in a gauge slice near
an instanton \(A_z\), let
\[
  L_{\Lambda,z}^{\rm bos}:N_z\mathcal M_\Lambda\longrightarrow
  N_z\mathcal M_\Lambda
\]
be the bosonic Hessian on the normal modes to the collective-coordinate
manifold, and let
\[
  M_{\Lambda,z}^{\rm gh}:C_{\Lambda,z}^{\rm gh}\longrightarrow
  C_{\Lambda,z}^{\rm gh}
\]
be the finite Faddeev--Popov ghost operator with residual stabilizer zero
modes removed.  For a fermion multiplet \(R\), write
\[
  \mathcal D_{\Lambda,z}^{R}:\mathcal S_{\Lambda,z}^{R,+}
  \longrightarrow \mathcal S_{\Lambda,z}^{R,-}
\]
for the regulated chiral Dirac operator.  Its zero modes are not included in
the nonzero-mode determinant; they are the Grassmann collective coordinates
whose saturation is described below.  The nonzero-mode determinant factor is
\begin{equation}
  \mathcal W_{\Lambda}^{\rm nz}(z)
  :=
  \frac{\det{}' M_{\Lambda,z}^{\rm gh}}
       {\det{}' M_{\Lambda,0}^{\rm gh}}
  \left(
    \frac{\det L_{\Lambda,0}^{\rm bos}}
         {\det{}' L_{\Lambda,z}^{\rm bos}}
  \right)^{1/2}
  \bigotimes_R
  \frac{\operatorname{Pf}{}' \mathcal D_{\Lambda,z}^{R}}
       {\operatorname{Pf} \mathcal D_{\Lambda,0}^{R}} .
  \label{eq:finite-regulator-instanton-nonzero-determinant-factor}
\end{equation}
The reference subscript \(0\) denotes the corresponding finite-regulator
operator in the chosen topologically trivial background and in the same
normalization of fields and gauge-group density.  The primes mean that
collective-coordinate zero modes, residual gauge zero modes, and fermion zero
modes have been removed before taking the finite determinant or Pfaffian.

For a vectorlike Dirac pair, after choosing the standard conjugate pairing,
the fermion factor in
\eqref{eq:finite-regulator-instanton-nonzero-determinant-factor} is an
ordinary scalar.  For a genuinely chiral fermion it is a vector in the
appropriate determinant or Pfaffian line over the background \(A_z\); a
scalar appears only after pairing this vector with the fermion zero-mode
insertions and with any anomaly-inflow or determinant-line trivialization
specified by the theory.  This is the same determinant-line convention used
in Chapter~\ref{ch:curved-eta-invariants-global-anomalies}.  Thus an instanton
amplitude is not a number until the anomaly-line pairing and zero-mode
saturation have also been specified.

With this notation the finite-regulator one-instanton contribution in a
single smooth chart, for insertions whose fermion zero-mode and anomaly-line
pairings have been specified, has the finite chart expression
\begin{equation}
  \dd\mu_{\Lambda,Q=1}(z)
  =
  \exp[-S_\Lambda(A_z)]\,
  \mathcal W_{\Lambda}^{\rm nz}(z)\,
  \dd\nu_{\Lambda,\mathcal M}(z),
  \label{eq:finite-regulator-instanton-measure-density}
\end{equation}
where \(\dd\nu_{\Lambda,\mathcal M}\) is the zero-mode density from
\eqref{eq:collective-coordinate-bosonic-measure}, or equivalently the ADHM
quotient density of Definition~\ref{def:adhm-quotient-density} after matching
the ADHM coordinates to the regulated gauge-field slice.  Formula
\eqref{eq:finite-regulator-instanton-measure-density} is a finite-dimensional
statement.  The continuum expression is the assertion that the product of
\(\mathcal W_{\Lambda}^{\rm nz}\), the coupling coordinate in
\(S_\Lambda\), and the renormalized operator insertions has a limit in the
chosen scheme.

The finite Gaussian origin of
\eqref{eq:finite-regulator-instanton-nonzero-determinant-factor} is as
follows.  A real bosonic normal integral with positive quadratic form \(L\)
gives \((\det L)^{-1/2}\) up to the universal power of \(2\pi\) already
accounted for in the stationary-phase convention.  A complex ghost Berezin
integral gives \(\det M^{\rm gh}\).  A real fermionic quadratic form gives a
Pfaffian; a complex chiral determinant is the corresponding determinant-line
version of the same finite exterior-algebra construction.  Removing zero
modes before taking these determinants is not optional: the removed
directions are precisely the collective coordinates or the fermion variables
which must be integrated or saturated separately.

Finally, changing the finite-regulator Lagrangian by a local counterterm
\(C_\Lambda[A]\) multiplies
\eqref{eq:finite-regulator-instanton-measure-density} by
\(\exp[-C_\Lambda(A_z)+C_\Lambda(0)]\).  If the one-instanton chart reduces
this difference to
\[
  C_\Lambda(A_z)-C_\Lambda(0)
  =
  c_0(\Lambda)+c_1(\Lambda)\log(\mu\rho)
  +R_\Lambda(\rho,U),
\]
then \(c_0\) changes the scheme-dependent overall constant, \(c_1\) changes
the displayed power of \(\mu\rho\), and the remainder \(R_\Lambda\) is part
of the unresolved boundary and continuum-limit analysis.  Counterterms are
therefore not an after-the-fact correction to a pre-existing determinant
constant; they are part of the regulated path-integral Lagrangian whose
finite determinant factor is being defined.

\paragraph{Small-instanton boundary exponent.}
For the charge-one stratum the Uhlenbeck boundary is reached by
\(\rho\to0\) at fixed center \(a\) and orientation \(U\).  The local boundary
question is therefore a statement about the \(\rho\)-asymptotics of the
finite-regulator density after the nonzero-mode determinant and the
renormalized insertions have been paired.  Let \(\mathcal X\) denote the
insertions under consideration, including the fermion zero-mode factors
needed to obtain a scalar from the determinant line.  A \emph{one-instanton
boundary exponent estimate} for \(\mathcal X\) consists of numbers
\(\beta_{\mathcal X}\in\mathbb R\) and \(\delta>0\), and a locally bounded
function \(A_{\mathcal X}(a,U)\), such that in a fixed compact set of centers
and in a fixed orientation chart the renormalized finite-regulator chart
density has a limit of the form
\begin{equation}
  \dd\mu_{Q=1,\mathcal X}
  =
  C_{\mathcal X}(\mu)\,
  \rho^{\,b_0+\beta_{\mathcal X}-5}
  \bigl(A_{\mathcal X}(a,U)+O(\rho^\delta)\bigr)\,
  \dd^4a\,\dd\rho\,\dd\Omega_{N_c}.
  \label{eq:one-instanton-boundary-exponent-estimate}
\end{equation}
The \(b_0-5\) part is the universal scale and one-loop running content in
\eqref{eq:one-instanton-universal-density}.  The exponent
\(\beta_{\mathcal X}\) records the net small-size contribution of the
renormalized operator insertions, fermion zero-mode wavefunctions, masses or
sources used to saturate zero modes, and any residual determinant factor
which has a finite power-law boundary behavior.  Logarithms may be included
by allowing \(A_{\mathcal X}\) to be a polynomial in \(\log(\mu\rho)\); they
leave the power threshold below unchanged.

On this boundary chart the local integral over \(0<\rho<\epsilon\) is
controlled by
\[
  \int_0^\epsilon
  \rho^{\,b_0+\beta_{\mathcal X}-5}\,
  (\log(\mu\rho))^q\,\dd\rho,
  \qquad q\in\mathbb Z_{\ge0}.
\]
It is locally finite at \(\rho=0\) precisely when
\begin{equation}
  b_0+\beta_{\mathcal X}>4 .
  \label{eq:one-instanton-small-rho-integrability-threshold}
\end{equation}
At equality the logarithmic integral diverges, and below equality the power
integral diverges.  Thus asymptotic freedom, \(b_0>0\), is a statement about
the running coupling at short distance, while small-instanton local
integrability of a chosen observable requires the stronger
observable-dependent condition
\eqref{eq:one-instanton-small-rho-integrability-threshold}.  For the
vacuum-density power with no extra small-\(\rho\) insertion factor
\(\beta_{\mathcal X}=0\), the local boundary test is \(b_0>4\).  Different
correlators have different \(\beta_{\mathcal X}\), and center collisions
with local operator insertions introduce additional boundary faces which
must be analyzed separately from the pure \(\rho\to0\) face.

This boundary exponent estimate is weaker than a complete multi-instanton
measure theorem.  It isolates the analytic estimate needed near a single
collapsed charge.  Higher charge involves simultaneous collisions of several
instantons, partial Uhlenbeck strata
\(\mathcal M_{k-\ell,N_c}\times\operatorname{Sym}^{\ell}(\mathbb R^4)\),
and determinant behavior transverse to each stratum.

\begin{definition}[Uhlenbeck boundary-face integrability datum]
\label{def:uhlenbeck-boundary-face-integrability-datum}
Fix an instanton charge \(k\), a gauge group \(SU(N_c)\), and a scalar
insertion datum \(\mathcal X\) including fermion-zero-mode and anomaly-line
pairings.  A Uhlenbeck boundary-face integrability datum consists of:
\begin{enumerate}[label=(\roman*)]
  \item a bubbling charge \(1\le \ell\le k\), a residual smooth charge
  \(k-\ell\), and a chart on the open Uhlenbeck stratum
  \[
    \mathcal M_{k-\ell,N_c}^{\rm smooth}
    \times
    \operatorname{Conf}_{\ell}(\mathbb R^4)/S_\ell ,
  \]
  with the pointlike instanton centers kept away from each other and from
  the local insertions;
  \item transverse gluing scales \(0<\rho_\alpha<\epsilon\), angular variables
  \(\omega_\alpha\), and boundary exponents \(A_\alpha(\mathcal X)\) for
  the \(\ell\) collapsed charge units;
  \item if some pointlike centers approach each other or an insertion support,
  collision distances \(0<d_\beta<\epsilon\) and collision exponents
  \(G_\beta(\mathcal X)\);
  \item a finite-regulator density bound, uniform on compact subsets of the
  residual smooth stratum and angular variables, of the form
  \begin{equation}
    |\dd\mu_{k,\mathcal X}|
    \le
    C\,\dd\nu_{k-\ell}^{\rm sm}
    \prod_{\alpha=1}^{\ell}
      \rho_\alpha^{A_\alpha(\mathcal X)-1}
      (1+|\log\rho_\alpha|)^{q_\alpha}\,
      \dd\rho_\alpha\,\dd\omega_\alpha
    \prod_{\beta}
      d_\beta^{G_\beta(\mathcal X)-1}
      (1+|\log d_\beta|)^{s_\beta}\,
      \dd d_\beta\,\dd\zeta_\beta ,
    \label{eq:uhlenbeck-boundary-face-density-bound}
  \end{equation}
  with nonnegative integers \(q_\alpha,s_\beta\).
\end{enumerate}
The exponents \(A_\alpha\) and \(G_\beta\) are part of the analytic
instanton-measure problem.  Classical ADHM dimension counts identify the
boundary coordinates; determinant, insertion, and counterterm estimates
supply the powers in
\eqref{eq:uhlenbeck-boundary-face-density-bound}.
\end{definition}

\begin{proposition}[Uhlenbeck face dimension and local integrability budget]
\label{prop:uhlenbeck-face-integrability-budget}
At a smooth residual instanton and \(\ell\) distinct pointlike ideal
instantons, the open Uhlenbeck stratum
\[
  \mathcal M_{k-\ell,N_c}^{\rm smooth}
  \times
  \operatorname{Conf}_{\ell}(\mathbb R^4)/S_\ell
\]
has local real codimension
\begin{equation}
  4\ell(N_c-1)
  \label{eq:uhlenbeck-open-stratum-codimension}
\end{equation}
inside the smooth charge-\(k\) ADHM dimension count.  If the \(\ell\)
pointlike charges are grouped into \(r\) collision clusters, the diagonal
collision stratum contributes the additional codimension \(4(\ell-r)\).

For a boundary-face datum satisfying
\eqref{eq:uhlenbeck-boundary-face-density-bound}, the local absolute
integral over the displayed boundary variables is finite whenever
\begin{equation}
  A_\alpha(\mathcal X)>0
  \quad\text{for every }\alpha,
  \qquad
  G_\beta(\mathcal X)>0
  \quad\text{for every }\beta .
  \label{eq:uhlenbeck-face-integrability-conditions}
\end{equation}
If one displayed exponent is zero and the leading coefficient on that face
has nonzero integral over the remaining bounded variables, the corresponding
local integral has a logarithmic divergence.  If one displayed exponent is
negative under the same nonvanishing condition, the divergence is a power
divergence.
\end{proposition}

\begin{proof}
The smooth framed charge-\(k\) dimension from
Proposition~\ref{prop:adhm-quotient-dimension-count} is \(4kN_c\).  On the
open Uhlenbeck stratum with \(\ell\) distinct ideal points, the residual
smooth instanton part has dimension \(4(k-\ell)N_c\), and the unordered
configuration space of \(\ell\) distinct points in \(\mathbb R^4\) has local
dimension \(4\ell\).  The codimension is therefore
\[
  4kN_c-\bigl(4(k-\ell)N_c+4\ell\bigr)
  =
  4\ell(N_c-1),
\]
which proves \eqref{eq:uhlenbeck-open-stratum-codimension}.  If the
\(\ell\) ideal points lie on a diagonal with \(r\) distinct collision
clusters, the point-configuration dimension drops from \(4\ell\) to \(4r\),
so the extra diagonal codimension is \(4(\ell-r)\).

For the integrability statement, the compact residual and angular variables
contribute only a finite multiplicative constant in the stated chart.  Each
scale integral is reduced to
\[
  \int_0^\epsilon
  t^{A-1}(1+|\log t|)^q\,\dd t .
\]
With \(u=-\log t\), this is bounded at \(t=0\) exactly when \(A>0\).  At
\(A=0\) it is the integral of a nonzero polynomial in \(u\) over an infinite
ray, hence logarithmic in the cutoff scale.  For \(A<0\), the antiderivative
has a negative power of the cutoff scale.  Applying this one-dimensional
criterion to every \(\rho_\alpha\) and \(d_\beta\) proves the product
criterion and the two divergence alternatives.
\end{proof}

For the charge-one boundary chart, there are no point-collision variables
and
\[
  A_1(\mathcal X)=b_0+\beta_{\mathcal X}-4 .
\]
Thus Proposition~\ref{prop:uhlenbeck-face-integrability-budget} recovers
\eqref{eq:one-instanton-small-rho-integrability-threshold}.  For \(k>1\) it
also records the extra data which a multi-instanton measure calculation must
provide: every bubbling scale, every diagonal collision of ideal instantons,
and every collision of an ideal instanton center with an inserted operator
has its own exponent.  A formula for the smooth ADHM quotient density and a
one-loop determinant on the principal stratum become a continuum
multi-instanton contribution only after these face estimates are supplied in
one compatible regulator and renormalization scheme.

\subsection{Scale and Running-Coupling Content of the One-Instanton Density}
\label{subsec:one-instanton-density-rg-content}

The preceding finite-regulator formula separates the zero-mode Jacobian from
the determinant on nonzero fluctuations.  The continuum one-instanton density
contains three universal pieces which can be fixed before computing the
scheme-dependent determinant constant.  First, the moduli
\[
  a^\mu\in\mathbb R^4,\qquad \rho\in\mathbb R_{>0}
\]
are the center and size of the instanton, while the remaining
\(4N_c-5\) coordinates are dimensionless orientation coordinates on the
embedded \(SU(2)\) orbit.  The translation-size part of a scale-covariant
local density must therefore be
\[
  \frac{\dd^4a\,\dd\rho}{\rho^5}.
\]
Indeed \(a\) and \(\rho\) both have length dimension one, so
\(\dd^4a\,\dd\rho/\rho^5\) is dimensionless and is invariant under the
simultaneous dilation \(a\mapsto\lambda a\), \(\rho\mapsto\lambda\rho\).
The power \(\rho^{-5}\) is the four-dimensional center plus one-dimensional
scale volume element.  It is the product of the zero-mode Jacobian and the
nonzero-mode determinant after their regulator-dependent powers of \(\rho\)
have been combined.

Second, the number of bosonic zero modes is \(m=4N_c\).  In the
semiclassical expansion of an action \(S[A]/g_{\rm ht}^2\), replacing the
zero-mode amplitudes by collective coordinates contributes one factor
\(g_{\rm ht}^{-1}\) for each bosonic zero mode.  Equivalently, after absorbing
normalization constants independent of \(g_{\rm ht}\) into the orientation
measure and determinant convention, the zero-mode normalization supplies
\[
  g_{\rm ht}^{-4N_c}
  =
  \left(\frac{8\pi^2}{g_{\rm ht}^2}\right)^{2N_c}
  \times
  \hbox{constant}.
\]
This is the origin of the power multiplying the one-instanton density.  The
constant depends on the precise normalization of the path-integral density
and of the invariant volume on the orientation orbit.

\paragraph{One-loop running factor in the one-instanton density.}
Use the common half-trace coupling \(g_{\rm ht}(\mu)\) in the BPST
subsection, and suppose the one-instanton density is renormalized in a
mass-independent scheme through one loop.  Let
\begin{equation}
  \mu\frac{\dd g_{\rm ht}}{\dd\mu}
  =
  -\frac{b_0}{16\pi^2}g_{\rm ht}^3+O(g_{\rm ht}^5),
  \qquad
  b_0=\frac{11}{3}N_c-\frac23N_f .
  \label{eq:instanton-density-beta0-definition}
\end{equation}
Then the part of the one-instanton measure fixed by scale covariance,
zero-mode normalization, and one-loop RG invariance is
\begin{equation}
  \dd\mu_{1}^{\rm univ}
  =
  C_{\mathcal S,N_c,N_f}\,
  \frac{\dd^4a\,\dd\rho}{\rho^5}\,
  \dd\Omega_{N_c}\,
  \left(\frac{8\pi^2}{g_{\rm ht}^2(\mu)}\right)^{2N_c}
  (\mu\rho)^{b_0}
  \exp\!\left[-\frac{8\pi^2}{g_{\rm ht}^2(\mu)}+\ii\theta\right]
  \left(1+O(g_{\rm ht}^2)\right).
  \label{eq:one-instanton-universal-density}
\end{equation}
Here \(\dd\Omega_{N_c}\) is a chosen invariant density on the local
orientation orbit
\[
  SU(N_c)/(SU(N_c-2)\times U(1)),
\]
with the evident \(SU(2)/\mathbb Z_2\) specialization when \(N_c=2\).
The coefficient \(C_{\mathcal S,N_c,N_f}\) is a finite constant depending on
the regulator, subtraction convention, and normalization of
\(\dd\Omega_{N_c}\); obtaining it requires the full renormalized determinant
calculation in the chosen scheme.
The factors
\(\dd^4a\,\dd\rho/\rho^5\), \(\dd\Omega_{N_c}\), and
\((8\pi^2/g_{\rm ht}^2)^{2N_c}\) have just been fixed by dimension,
orientation invariance, and the \(4N_c\) bosonic zero modes.  It remains to
determine the power of \(\mu\rho\) at one loop.  Write the unknown power as
\(\alpha\) and apply \(\mu\dd/\dd\mu\) to the logarithm of the density factor
\[
  g_{\rm ht}^{-4N_c}
  (\mu\rho)^\alpha
  \exp\!\left[-\frac{8\pi^2}{g_{\rm ht}^2(\mu)}\right].
\]
The derivative is
\[
  \alpha
  -4N_c\,\frac{\beta(g_{\rm ht})}{g_{\rm ht}}
  +
  \frac{16\pi^2}{g_{\rm ht}^3}\,\beta(g_{\rm ht}) .
\]
Substituting \eqref{eq:instanton-density-beta0-definition} gives
\[
  \alpha-b_0+O(g_{\rm ht}^2).
\]
RG invariance of a physical one-instanton contribution through one loop
therefore fixes \(\alpha=b_0\).  The derivative of the zero-mode power
starts at order \(g_{\rm ht}^2\), so it belongs to the next perturbative order
and is absorbed into the higher-loop correction multiplying
\eqref{eq:one-instanton-universal-density}.  The finite constant
\(C_{\mathcal S,N_c,N_f}\) is outside this argument because RG invariance
fixes scale dependence, while the constant term is set by the actual
regularized determinant and measure normalization.

\begin{proposition}[Pure-gauge Pauli--Villars determinant constant]
\label{prop:pure-gauge-pv-instanton-determinant-constant}
In pure \(SU(N)\) Yang--Mills, \(N\geq2\), use the common half-trace
coupling \(g_{\rm PV}(\mu)\) and the Pauli--Villars determinant convention of
the one-instanton calculation
\cite{tHooft1976Pseudoparticle,Bernard1979InstantonDeterminants}.  For
gauge-invariant insertions, integrate the embedded orientation orbit with the
corresponding Haar normalization.  Then the one-loop, one-instanton density is
\begin{equation}
  \dd n^{\rm PV}_{I,N}
  =
  \frac{\dd^4a\,\dd\rho}{\rho^5}\,
  C^{\rm PV}_{N}\,
  \left(\frac{8\pi^2}{g_{\rm PV}^2(\mu)}\right)^{2N}
  (\mu\rho)^{11N/3}
  \exp\!\left[-\frac{8\pi^2}{g_{\rm PV}^2(\mu)}+\ii\theta\right]
  \left(1+O(g_{\rm PV}^2)\right),
  \label{eq:pure-gauge-pv-instanton-density}
\end{equation}
where the finite constant is
\begin{equation}
  C^{\rm PV}_{N}
  =
  \frac{2^{2-2N}}{\pi^2 (N-1)!(N-2)!}\,
  \exp\!\left[-\alpha(1)-2(N-2)\alpha\!\left(\frac12\right)\right].
  \label{eq:pure-gauge-pv-instanton-constant-alpha}
\end{equation}
Here
\[
  \alpha(1)=0.443307\ldots,\qquad
  \alpha\!\left(\frac12\right)=0.145873\ldots .
\]
Equivalently, in the conventional rounded form,
\begin{equation}
  C^{\rm PV}_{N}
  =
  \frac{0.466\,\exp[-1.679\,N]}{(N-1)!(N-2)!}
  \quad\hbox{to the displayed precision}.
  \label{eq:pure-gauge-pv-instanton-constant-rounded}
\end{equation}
For example \(C^{\rm PV}_{3}\simeq1.51\times10^{-3}\).
\end{proposition}

\begin{proof}
The orientation-integrated singular-gauge collective-coordinate calculation
gives, after the Pauli--Villars zero-mode mass factors have been combined with
the logarithmic determinant into \((\mu\rho)^{11N/3}\), the residual
zero-mode coefficient
\[
  \frac{2^{4N+2}\pi^{4N-2}}{(N-1)!(N-2)!}\,g_{\rm PV}^{-4N}.
\]
Rewriting
\[
  g_{\rm PV}^{-4N}
  =
  \left(\frac{8\pi^2}{g_{\rm PV}^2}\right)^{2N}
  (8\pi^2)^{-2N}
\]
turns this residual coefficient into
\[
  \frac{2^{2-2N}}{\pi^2 (N-1)!(N-2)!}
  \left(\frac{8\pi^2}{g_{\rm PV}^2}\right)^{2N}.
\]
The finite part of the nonzero-mode determinant contributes
\[
  \exp\!\left[-\alpha(1)-2(N-2)\alpha\!\left(\frac12\right)\right],
\]
where \(\alpha(1)\) is the \(SU(2)\) vector-ghost block and the
\(2(N-2)\) copies of \(\alpha(1/2)\) are the additional doublet blocks in the
embedded \(SU(2)\subset SU(N)\) decomposition.  Multiplying these two factors
gives \eqref{eq:pure-gauge-pv-instanton-constant-alpha}; combining it with
the classical action and logarithmic determinant gives
\eqref{eq:pure-gauge-pv-instanton-density}.  Taking logarithms of
\eqref{eq:pure-gauge-pv-instanton-constant-alpha} yields
\[
  \log C^{\rm PV}_{N}
  =
  \log(0.466\ldots)-1.679\ldots\,N
  -\log((N-1)!)-\log((N-2)!),
\]
which is the rounded form
\eqref{eq:pure-gauge-pv-instanton-constant-rounded}.
\end{proof}

\begin{proposition}[Heat-kernel source of the instanton logarithm]
\label{prop:instanton-heat-kernel-beta0-logarithm}
In the same background-field gauge and half-trace convention, the logarithmic
part of the charge-\(k\) nonzero-mode fluctuation determinant has
\begin{equation}
  \partial_{\log\mu}\log\mathcal W_{\rm log}^{\rm nz}
  =
  k\left[
    \frac{11}{3}C_2(G)
    -\frac{4}{3}\sum_{\text{\rm Dirac }R}T(R)
  \right].
  \label{eq:instanton-heat-kernel-beta0-logarithm}
\end{equation}
For vectorlike \(SU(N_c)\) QCD with \(N_f\) Dirac fundamentals this coefficient
is \(k b_0\), with \(b_0\) as in
\eqref{eq:instanton-density-beta0-definition}.  Hence, in a charge-one BPST
chart,
\begin{equation}
  \log\mathcal W_{\rm log}^{\rm nz}(a,\rho,U;\mu)
  =
  b_0\log(\mu\rho)+\log\mathcal W_{\rm fin}^{\rm nz}(U;\mathcal S),
  \label{eq:instanton-nonzero-mode-log-mu-rho}
\end{equation}
up to higher-loop and local finite-subtraction terms in the chosen scheme.
Equivalently, if \(X_\kappa=8\pi^2/g_{\rm ht}^2(\kappa)\) and
\(X_\Lambda=X_\mu+b_0\log(\Lambda/\mu)\) at one loop, then
\begin{equation}
  \exp[-X_\Lambda+\ii\theta]\,
  (\Lambda\rho)^{b_0}
  =
  \exp[-X_\mu+\ii\theta]\,
  (\mu\rho)^{b_0}.
  \label{eq:instanton-cutoff-renormalized-scale-trade}
\end{equation}
The factor \((\mu\rho)^{b_0}\) in
\eqref{eq:one-instanton-universal-density} is therefore the logarithmic
shadow of the nonzero-mode fluctuation spectrum plus coupling
renormalization, while the finite constant still requires the full spectral
determinant calculation.
\end{proposition}

\begin{proof}
The small-proper-time expansion of
\eqref{eq:proper-time-instanton-determinant-ratio} is local in the gauge
background.  With background-field gauge fixing, the vector fluctuation and
ghost complex contribute the coefficient \((11/3)C_2(G)\) multiplying the
topological charge in the logarithmic counterterm, and a Dirac fermion in
representation \(R\) contributes \(-(4/3)T(R)\).  Summing the fields gives
\eqref{eq:instanton-heat-kernel-beta0-logarithm}.  For a fundamental Dirac
fermion of \(SU(N_c)\), \(T(\square)=1/2\), so the matter term is
\(-2N_f/3\).  The coefficient in this convention is therefore
\[
  \frac{11}{3}N_c-\frac{2}{3}N_f=b_0 .
\]

A BPST connection with size \(\rho\) is obtained from the unit-size connection
by translation, global orientation, and dilation.  After the zero modes are
removed, the logarithmic scale dependence of the determinant can therefore
depend on the size only through the dimensionless combination \(\mu\rho\),
apart from finite scheme-dependent terms.  This gives
\eqref{eq:instanton-nonzero-mode-log-mu-rho}.  Finally, the one-loop running of
\(X_\kappa\) gives \(X_\Lambda=X_\mu+b_0\log(\Lambda/\mu)\); substituting this
identity into the cutoff expression proves
\eqref{eq:instanton-cutoff-renormalized-scale-trade}.
\end{proof}

Fermion zero modes are controlled by the index theorem already stated in
\eqref{eq:dirac-index-anomaly}.  In the conventional half-trace notation, a
left-handed Weyl fermion in representation \(R\) has
\begin{equation}
  n_+-n_-=2T_R\,k,
  \qquad
  \operatorname{tr}_R(T_R^aT_R^b)=T_R\delta^{ab}.
  \label{eq:instanton-index-dynkin}
\end{equation}
For a fundamental Weyl fermion of \(SU(N_c)\), \(T_\square=1/2\), so a
\(k=1\) instanton has one normalizable zero mode of the chirality selected by
the self-dual background.  The anti-instanton has the opposite chirality.  In
the trace-delta convention the numerical value called \(T_R\) is doubled, but
the topological density and the coupling coordinate are correspondingly
rewritten; the integer kernel dimension in
\eqref{eq:instanton-index-dynkin} is the invariant statement.

The zero modes determine the Grassmann selection rule.  Expand a massless
fermion in an instanton background as
\[
  \psi(x)=\sum_{\alpha=1}^{n_0}\xi_\alpha\psi_{0,\alpha}(x)
  +\sum_{\lambda\neq0}\xi_\lambda\psi_\lambda(x),
\]
where \(\psi_{0,\alpha}\) are the zero modes and \(\xi_\alpha\) are odd
coefficients.  The quadratic fermion action contains no term involving
\(\xi_\alpha\).  Therefore the Berezin integral over zero-mode variables
vanishes unless operator insertions supply every missing \(\xi_\alpha\).
This is the finite-dimensional Berezin rule \(\int\dd\xi\,1=0\),
\(\int\dd\xi\,\xi=1\) applied after the spectral decomposition of the
regulated fermion field.

\begin{proposition}[Instanton sectors as physical correlator contributions]
\label{prop:instanton-sector-physical-correlator-contribution}
Fix a finite regulator, a \(Q=1\) instanton chart with collective coordinate
\(z=(a,\rho,U)\), smooth chart domain \(\mathcal U_{Q=1}\), and
renormalized insertions \(\mathcal O_1,\ldots,\mathcal O_r\).  After gauge
fixing, collective coordinate separation, and integration of nonzero
fluctuations, the instanton sector contribution has the form
\begin{equation}
  \bigl\langle\mathcal O_1\cdots\mathcal O_r\bigr\rangle_{Q=1,\Lambda}
  =
  \int_{\mathcal U_{Q=1}}
  \ee^{-S_\Lambda(A_z)+\ii\theta}\,
  \mathcal W_{\Lambda}^{\rm nz}(z)\,
  \mathcal Z_{\mathcal O,\Lambda}^{0}(z)\,
  \dd\nu_{\Lambda,\mathcal M}(z),
  \label{eq:instanton-physical-correlator-assembly}
\end{equation}
provided the determinant/Pfaffian line has been paired with the fermion
zero-mode and anomaly-line data to produce a scalar density.  Here
\(\mathcal W_{\Lambda}^{\rm nz}\) is the nonzero-mode determinant datum from
\eqref{eq:finite-regulator-instanton-nonzero-determinant-factor}, while
\(\mathcal Z_{\mathcal O,\Lambda}^{0}(z)\) is the finite Berezin coefficient
\begin{equation}
  \mathcal Z_{\mathcal O,\Lambda}^{0}(z)
  =
  \int\dd\xi_{n_0}\cdots\dd\xi_1\,
  \mathcal O_{1,\Lambda}^{0}(z,\xi)\cdots
  \mathcal O_{r,\Lambda}^{0}(z,\xi)\,
  \exp[-S_{\rm lift,\Lambda}^{0}(z,\xi)] .
  \label{eq:instanton-zero-mode-saturation-functional}
\end{equation}
The superscript \(0\) means that each insertion and each mass, Yukawa, or
source term has been restricted to the zero-mode subspace.  If the integrand
in \eqref{eq:instanton-zero-mode-saturation-functional} does not contain the
monomial \(\xi_1\cdots\xi_{n_0}\), then
\(\mathcal Z_{\mathcal O,\Lambda}^{0}(z)=0\).  If it does contain that
monomial, the coefficient of that monomial is the only part of the insertions
which is seen by the instanton sector.
\end{proposition}

\begin{proof}
At finite cutoff the gauge-fixed fields split into collective coordinates,
normal bosonic modes, ghosts, fermion zero modes, and fermion nonzero modes.
Gaussian integration over the normal bosonic and ghost variables gives the
determinant factors already isolated in
\(\mathcal W_{\Lambda}^{\rm nz}\); Gaussian or Pfaffian integration over
fermion nonzero modes gives the remaining nonzero-mode determinant-line
factor.  None of these operations integrates over the zero-mode variables
\(\xi_\alpha\), because the quadratic fermion operator vanishes on them.
The only finite-dimensional integral left in the fermionic zero-mode sector is
\eqref{eq:instanton-zero-mode-saturation-functional}.  Berezin integration
extracts the coefficient of the ordered top monomial
\(\xi_1\cdots\xi_{n_0}\).  All terms missing at least one zero-mode variable
therefore vanish, while terms containing the top monomial contribute exactly
their coefficient.  Multiplying this coefficient by the bosonic
collective-coordinate density and the nonzero-mode determinant gives
\eqref{eq:instanton-physical-correlator-assembly}.
\end{proof}

Thus the moduli space is not the physical answer; it is the domain of a
semiclassical amplitude.  The physical question is which insertions, masses,
or sources saturate the fermion zero modes, what nonzero-mode determinant and
counterterm scheme define \(\mathcal W_{\Lambda}^{\rm nz}\), and whether the
resulting \(a,\rho,U\) integral is finite at its boundary faces.

\paragraph{From collective coordinates to an amplitude.}
For later use, it is helpful to compress the \(Q=1\) calculation into the
ordered finite-regulator pipeline
\begin{equation}
  K_{\Lambda,\mathcal J}^{E}(z)
  =
  \ee^{-S_\Lambda(A_z)+\ii\theta}\,
  \mathcal W_{\Lambda}^{\rm nz}(z)\,
  \mathcal Z_{\mathcal J,\Lambda}^{0}(z)\,
  \dd\nu_{\Lambda,\mathcal M}(z),
  \qquad
  G_{\Lambda,\mathcal J}^{E,Q=1}
  =
  \int_{\mathcal U_{Q=1}}K_{\Lambda,\mathcal J}^{E}(z).
  \label{eq:instanton-amplitude-pipeline-euclidean-kernel}
\end{equation}
The first factor selects the topological sector and its classical action; the
last factor is the collective-coordinate density; the middle two factors are
where the fluctuation and fermion-saturation calculation enters.  A
semiclassical local vertex is obtained only after replacing
\(K_{\Lambda,\mathcal J}^{E}\) on a declared window by a leading kernel and
then projecting it onto an observable coordinate:
\begin{equation}
  \mathcal A_{\rm phys}^{Q=1}[\mathcal J]
  =
  \operatorname*{Proj}_{\rm phys}
  \operatorname*{Cont}_{\mathcal S}
  \int_{\mathcal U_{\rm win}}
  K_{{\rm lead},\mathcal J}^{E}(z)
  +
  R_{\rm det}+R_{\rm zm}+R_{\rm src}+R_{\rm Schur}
  +R_{\rho}+R_{\rm cont}+R_{\rm spec}+R_{\rm IR}+R_{\rm cut}.
  \label{eq:instanton-amplitude-pipeline-physical-map}
\end{equation}
Here the determinant residual includes the nonzero-mode spectrum and local
counterterms; \(R_{\rm zm}\) records the zero-mode projection or derivative
expansion of the sources; \(R_{\rm src}\) records the color-singlet,
hadronic, or OPE matching layer; \(R_{\rm Schur}\) records finite
nonzero-mode corrections to the zero-mode block; \(R_{\rho}\) is the size and
Uhlenbeck-face endpoint remainder; and the final four terms are the
continuation, spectral-projection, infrared, and unitarity-cut remnants
needed before a physical amplitude is asserted.

This ordering is part of the content of the instanton calculation.  The ADHM
or BPST moduli space supplies the domain and the collective-coordinate
Jacobian.  It does not supply the nonzero-mode determinant, the fermion
zero-mode Berezin coefficient, the source-to-observable matching, or the
endpoint estimates.  Conversely, the local 't Hooft operator is the
small-size, source-adiabatic coordinate of
\eqref{eq:instanton-amplitude-pipeline-euclidean-kernel}; the four-point
amplitude calculation keeps the zero-mode form factors, determinant
normalization, external source residues, and size tails inside the same
integral until a hard, screened, thermal, or Wilsonian window justifies
separating them.

For vectorlike QCD with \(N_f\) massless Dirac fundamentals, a \(Q=1\)
instanton supplies one zero mode for each flavor in one chirality sector.  The
semiclassical functional integral therefore generates the following leading
local term in the derivative expansion around external fields whose variation
scale is much larger than \(\rho\):
\begin{equation}
  \mathcal L_{\rm inst}
  =
  \int_0^\infty\frac{\dd\rho}{\rho^5}\,
  C_{N_c,N_f}(\rho,\mu)
  \exp\!\left(-\frac{8\pi^2}{g_{\rm ht}^2(\mu)}+\ii\theta\right)
  \det_{1\le f,f'\le N_f}
  \left[
    \rho^3\bar\psi_{R f}\psi_{L f'}
  \right]
  +\hbox{anti-instanton conjugate}
  \label{eq:qcd-thooft-vertex-leading-local}
\end{equation}
after integrating over the instanton center and color orientation and after
contracting spin and color indices as dictated by the zero-mode wavefunctions
and gauge invariance.  The determinant is over flavor indices; it is the local
tensor compatible with the nonabelian chiral flavor symmetry.  For flavor
diagonal external sources it reduces to the corresponding alternating sum of
products of \(N_f\) bilinears.  At one loop
\[
  C_{N_c,N_f}(\rho,\mu)
  =
  C^{(0)}_{N_c,N_f}
  \left(\frac{8\pi^2}{g_{\rm ht}^2(\mu)}\right)^{2N_c}
  (\rho\mu)^{b_0}
  \left(1+O(g_{\rm ht}^2(\mu))\right),
  \qquad
  b_0=\frac{11}{3}N_c-\frac23N_f,
\]
where \(b_0\) is defined by
\(\mu\,\dd g_{\rm ht}/\dd\mu=-b_0g_{\rm ht}^3/(16\pi^2)+O(g_{\rm ht}^5)\).
The scale and coupling dependence is the universal content of the running-factor
calculation above; the constant
\(C^{(0)}_{N_c,N_f}\) is fixed only after choosing the determinant and
orientation-volume normalization.  The dimensionless factors
\(\rho^3\bar\psi_R\psi_L\) in
\eqref{eq:qcd-thooft-vertex-leading-local} record the normalization of the
external zero-mode wavefunctions in the local limit; changing the
regularization scheme changes \(C^{(0)}_{N_c,N_f}\) and higher-loop factors,
not the zero-mode tensor or the selection rule.  When the external fields vary
on the scale \(\rho\), the same zero-mode integration gives a nonlocal kernel
rather than the local determinant in
\eqref{eq:qcd-thooft-vertex-leading-local}.  Its regulator-independent
selection rule is
\[
  \Delta Q_A=2N_f
\]
for the anomalous axial charge, in agreement with the integrated anomaly.
Small quark masses may replace missing fermion insertions by mass factors,
because the mass term pairs the corresponding zero-mode variables.

\begin{proposition}[Flavor determinant from the instanton zero modes]
\label{prop:qcd-thooft-determinant-zero-mode-amplitude}
In the \(Q=1\) sector of vectorlike \(SU(N_c)\) QCD, let
\(\lambda_f,\widetilde\lambda_f\), \(f=1,\ldots,N_f\), denote the two
Grassmann zero-mode coefficients paired by the local source matrix
\(B=(B_{ff'})\).  Choose the zero-mode orientation so that
\begin{equation}
  \int
  \prod_{f=N_f}^{1}\dd\widetilde\lambda_f\,\dd\lambda_f\,
  \exp\!\left(
    \sum_{f,f'=1}^{N_f}
    \widetilde\lambda_f B_{ff'}\lambda_{f'}
  \right)
  =
  \det B .
  \label{eq:thooft-vertex-berezin-determinant}
\end{equation}
For slowly varying external quark fields,
\(B_{ff'}=\rho^3\bar\psi_{Rf}\psi_{Lf'}\), and
\eqref{eq:thooft-vertex-berezin-determinant} is precisely the flavor
determinant in \eqref{eq:qcd-thooft-vertex-leading-local}.  If instead all
zero modes are saturated by the quark mass term, the same finite-dimensional
calculation replaces \(B\) by the mass-overlap matrix
\(B^{(m)}_{ff'}(z)\).  Its phase is the phase of \(\det M\), so the
one-instanton vacuum term has phase
\(\exp[\ii(\theta+\arg\det M)]\), up to the real positive overlap and
determinant normalizations.
\end{proposition}

\begin{proof}
Expand the exponential in
\eqref{eq:thooft-vertex-berezin-determinant}.  Every contributing top-degree
term must contain each \(\widetilde\lambda_f\) once and each \(\lambda_f\)
once.  Such a term is specified by a permutation \(\sigma\in S_{N_f}\):
\[
  \prod_{f=1}^{N_f}
  \widetilde\lambda_f B_{f,\sigma(f)}\lambda_{\sigma(f)} .
\]
Reordering the \(\lambda_{\sigma(f)}\) variables into the chosen Berezin
orientation gives the sign \(\operatorname{sgn}\sigma\).  Berezin integration
therefore extracts
\[
  \sum_{\sigma\in S_{N_f}}
  \operatorname{sgn}\sigma
  \prod_{f=1}^{N_f}B_{f,\sigma(f)}
  =
  \det B .
\]
For \(B_{ff'}=\rho^3\bar\psi_{Rf}\psi_{Lf'}\), this gives the local
\(2N_f\)-fermion operator.  If \(B\) is supplied by the mass term, each
zero-mode overlap is real after the same regulator and zero-mode orientation
have been fixed, so the only flavor phase is \(\arg\det M\).
\end{proof}

\begin{proposition}[Two-flavor channel content of the local instanton vertex]
\label{prop:two-flavor-thooft-channel-decomposition}
For \(N_f=2\), set
\[
  \Phi_{ff'}=\bar q_{Rf}q_{Lf'},\qquad
  S_{ff'}=\bar q_fq_{f'},\qquad
  P_{ff'}=\bar q_f\ii\gamma_5q_{f'} .
\]
With the chiral projectors used in this chapter,
\begin{equation}
  \Phi=\frac12(S+\ii P).
  \label{eq:two-flavor-chiral-bilinear-matrix}
\end{equation}
Write the scalar and pseudoscalar flavor matrices in the Pauli basis as
\[
  S=s_0\mathbf 1+\vec s\cdot\vec\tau,\qquad
  P=p_0\mathbf 1+\vec p\cdot\vec\tau .
\]
Then the local instanton plus anti-instanton flavor factor is
\begin{equation}
\begin{aligned}
  \ee^{\ii\theta}\det\Phi
  +\ee^{-\ii\theta}\det\Phi^\dagger
  &=
  \frac{\cos\theta}{2}
  \left(s_0^2-\vec s^{\,2}-p_0^2+\vec p^{\,2}\right) \\
  &\qquad
  -\sin\theta\left(s_0p_0-\vec s\cdot\vec p\right).
\end{aligned}
  \label{eq:two-flavor-thooft-channel-decomposition}
\end{equation}
At \(\theta=0\), the one-instanton local vertex therefore separates the
isosinglet scalar, isotriplet scalar, isosinglet pseudoscalar, and isotriplet
pseudoscalar channels as
\begin{equation}
  \det\Phi+\det\Phi^\dagger
  =
  \frac12
  \left(s_0^2-\vec s^{\,2}-p_0^2+\vec p^{\,2}\right).
  \label{eq:two-flavor-thooft-theta-zero-channels}
\end{equation}
The determinant is invariant under
\(\Phi\mapsto V_R^\dagger\Phi V_L\), \(V_L,V_R\in SU(2)\), but under the
flavor-singlet axial rotation \(q\mapsto\ee^{\ii\alpha\gamma_5}q\) it carries
phase \(\det\Phi\mapsto\ee^{-4\ii\alpha}\det\Phi\).  Hence
\(\ee^{\ii\theta}\det\Phi+\hbox{\rm conjugate}\) is the local channel form of
the same anomalous \(U(1)_A\) selection rule as
\eqref{eq:qcd-thooft-vertex-leading-local}: a compensating
\(\theta\mapsto\theta+4\alpha\) keeps the \(Q=1\) contribution in the same
coordinate.

This channel decomposition is only the zeroth local term of the source
convolution.  The coefficient multiplying
\eqref{eq:two-flavor-thooft-channel-decomposition} is still the
size/orientation/nonzero-mode determinant integral discussed below.  The formula
therefore explains which chiral channels the zero modes select; it is not a
standalone prediction of scalar or pseudoscalar bound-state masses.
\end{proposition}

\begin{proof}
The projector identity gives
\[
  \bar q_Rq_L=\bar qP_Lq
  =\frac12(\bar qq-\bar q\gamma_5q)
  =\frac12(\bar qq+\ii\,\bar q\ii\gamma_5q),
\]
which is \eqref{eq:two-flavor-chiral-bilinear-matrix}.  For any matrix
\(X=x_0\mathbf 1+\vec x\cdot\vec\tau\), the Pauli relation
\(\tau_a\tau_b=\delta_{ab}\mathbf 1+\ii\epsilon_{abc}\tau_c\) gives
\(\det X=x_0^2-\vec x^{\,2}\).  Applying this to
\(X=S+\ii P\) gives
\[
  \det(S+\ii P)
  =
  s_0^2-\vec s^{\,2}-p_0^2+\vec p^{\,2}
  +2\ii(s_0p_0-\vec s\cdot\vec p).
\]
Since \(\det\Phi=\frac14\det(S+\ii P)\), taking
\(\ee^{\ii\theta}\det\Phi+\ee^{-\ii\theta}\det\Phi^\dagger\) proves
\eqref{eq:two-flavor-thooft-channel-decomposition}.  The
\(\theta=0\) specialization is immediate.

Finally, an \(SU(2)_L\times SU(2)_R\) rotation multiplies \(\Phi\) by
unit-determinant matrices on the two sides, so its determinant is unchanged.
For \(q\mapsto\ee^{\ii\alpha\gamma_5}q\), one has
\(\bar q_Rq_L\mapsto\ee^{-2\ii\alpha}\bar q_Rq_L\), hence
\(\det\Phi\mapsto\ee^{-4\ii\alpha}\det\Phi\).
\end{proof}

\begin{controlledapproximation}[Instanton contribution to two-flavor chiral
source curvatures]
\label{ca:two-flavor-instanton-source-susceptibility}
Suppose the finite-regulator \(Q=\pm1\) calculation, including the
size/orientation integral, nonzero-mode determinant, source normalization, and
local derivative expansion, supplies a finite coefficient
\(\kappa_{\rm inst}^{(2)}(\mu)\) multiplying the local \(N_f=2\) source density
of Proposition~\ref{prop:two-flavor-thooft-channel-decomposition}.  Equivalently,
for constant renormalized scalar and pseudoscalar source coordinates
\[
  S=s_0\mathbf 1+\vec s\cdot\vec\tau,\qquad
  P=p_0\mathbf 1+\vec p\cdot\vec\tau ,
\]
the instanton part of the source generating-density coordinate
\(\mathcal W=V_4^{-1}\log Z\) has quadratic term
\begin{equation}
\begin{aligned}
  \mathcal W_{\rm inst}^{(2)}
  &=
  \kappa_{\rm inst}^{(2)}
  \left[
    \frac{\cos\theta}{2}
    \left(s_0^2-\vec s^{\,2}-p_0^2+\vec p^{\,2}\right)
    -\sin\theta\left(s_0p_0-\vec s\cdot\vec p\right)
  \right]  \\
  &\qquad
  +O(J^3,\partial J),
\end{aligned}
  \label{eq:two-flavor-instanton-source-curvature-functional}
\end{equation}
where \(J\) denotes the collection of source coordinates.  Define the local
source curvatures by differentiating \(\mathcal W\) at zero source:
\[
  \chi_\sigma=\partial_{s_0}^2\mathcal W,\quad
  \chi_{\delta_a}=\partial_{s_a}^2\mathcal W,\quad
  \chi_\eta=\partial_{p_0}^2\mathcal W,\quad
  \chi_{\pi_a}=\partial_{p_a}^2\mathcal W,
  \qquad\hbox{no sum on }a .
\]
Then the local instanton contribution is
\begin{equation}
  \chi_\sigma^{\rm inst}
  =
  \chi_{\pi_a}^{\rm inst}
  =
  \kappa_{\rm inst}^{(2)}\cos\theta,
  \qquad
  \chi_{\delta_a}^{\rm inst}
  =
  \chi_{\eta}^{\rm inst}
  =
  -\kappa_{\rm inst}^{(2)}\cos\theta ,
  \label{eq:two-flavor-instanton-channel-curvatures}
\end{equation}
and the CP-odd source-mixing curvatures are
\begin{equation}
  \partial_{s_0}\partial_{p_0}\mathcal W_{\rm inst}^{(2)}
  =
  -\kappa_{\rm inst}^{(2)}\sin\theta,
  \qquad
  \partial_{s_a}\partial_{p_b}\mathcal W_{\rm inst}^{(2)}
  =
  \kappa_{\rm inst}^{(2)}\sin\theta\,\delta_{ab}.
  \label{eq:two-flavor-instanton-cp-odd-source-mixing}
\end{equation}
At \(\theta=0\), the two \(U(1)_A\)-paired differences supplied by this
finite-activity term are therefore
\begin{equation}
  \left(\chi_{\pi_a}-\chi_{\delta_a}\right)_{\rm inst}
  =
  \left(\chi_\sigma-\chi_\eta\right)_{\rm inst}
  =
  2\kappa_{\rm inst}^{(2)} .
  \label{eq:two-flavor-instanton-u1a-susceptibility-splitting}
\end{equation}

Equations
\eqref{eq:two-flavor-instanton-channel-curvatures}--\eqref{eq:two-flavor-instanton-u1a-susceptibility-splitting}
are source-curvature statements in the declared local window.  The full
zero-momentum QCD susceptibilities also contain connected long-distance
propagation, Goldstone poles in the chiral limit, ordinary perturbative and
confining contributions, and finite contact-term conventions.  Thus the
instanton calculation identifies the \(U(1)_A\)-violating local curvature it
adds to the generating functional.  A spectrum statement additionally requires
the color-singlet spectral representation, long-distance matrix elements, and
the continuum/volume limits in the chosen QCD phase.
\end{controlledapproximation}

\begin{proposition}[Axial Ward ledger for the instanton source curvatures]
\label{prop:two-flavor-instanton-source-ward-ledger}
The local two-flavor instanton source functional
\eqref{eq:two-flavor-instanton-source-curvature-functional} obeys the
finite-dimensional anomalous singlet-axial Ward identity
\begin{equation}
  \left[
    4\partial_\theta
    +2p_0\partial_{s_0}
    +2\sum_{a=1}^{3}p_a\partial_{s_a}
    -2s_0\partial_{p_0}
    -2\sum_{a=1}^{3}s_a\partial_{p_a}
  \right]
  \mathcal W_{\rm inst}^{(2)}
  =
  O(J^3,\partial J).
  \label{eq:two-flavor-instanton-source-ward-identity}
\end{equation}
Here the source coordinates transform as the mass/source spurions in the
finite-regulator path integral:
\[
  \delta_\alpha S=2\alpha P,\qquad
  \delta_\alpha P=-2\alpha S,\qquad
  \delta_\alpha\theta=4\alpha .
\]
Thus the same anomalous Jacobian which shifts \(\theta\) also rotates the
local scalar and pseudoscalar source coordinates.  The susceptibility
splittings in
\eqref{eq:two-flavor-instanton-u1a-susceptibility-splitting} are the
second-source derivatives of the same anomalous Ward identity which produces
the 't Hooft vertex.

More explicitly, for each paired coordinate \(s,p\), with \(s=s_0,p=p_0\)
or \(s=s_a,p=p_a\), define
\[
  \chi_s=\partial_s^2\mathcal W_{\rm inst}^{(2)},\qquad
  \chi_p=\partial_p^2\mathcal W_{\rm inst}^{(2)},\qquad
  m_{sp}=\partial_s\partial_p\mathcal W_{\rm inst}^{(2)}
\]
at zero source.  Then the differentiated Ward identity gives
\begin{equation}
  \partial_\theta\chi_s=m_{sp},\qquad
  \partial_\theta\chi_p=-m_{sp},\qquad
  \chi_s-\chi_p=-2\,\partial_\theta m_{sp}.
  \label{eq:two-flavor-instanton-curvature-ward-relations}
\end{equation}
For the singlet pair this is the \((\sigma,\eta)\) relation; for a triplet
pair it is the \((\delta_a,\pi_a)\) relation, so reversing the order of the
difference gives the positive \(\pi_a-\delta_a\) splitting displayed above.
\end{proposition}

\begin{proof}
Under an infinitesimal singlet axial rotation,
\(S+\ii P\mapsto (1-2\ii\alpha)(S+\ii P)\), which gives
\(\delta S=2\alpha P\) and \(\delta P=-2\alpha S\).  The anomaly shifts the
source-sector theta coordinate by \(\delta\theta=4\alpha\) for \(N_f=2\).
Consequently the infinitesimal Ward vector field is the operator displayed in
\eqref{eq:two-flavor-instanton-source-ward-identity}.

Let
\[
  A=s_0^2-\vec s^{\,2}-p_0^2+\vec p^{\,2},
  \qquad
  B=s_0p_0-\vec s\cdot\vec p .
\]
The source rotation gives
\[
  \delta_\alpha A=8\alpha B,\qquad
  \delta_\alpha B=-2\alpha A .
\]
The theta derivative of the quadratic part of
\eqref{eq:two-flavor-instanton-source-curvature-functional} contributes
\[
  4\alpha\,\kappa_{\rm inst}^{(2)}
  \left[-\frac{\sin\theta}{2}A-\cos\theta\,B\right],
\]
whereas the source rotation contributes
\[
  \kappa_{\rm inst}^{(2)}
  \left[4\alpha\cos\theta\,B+2\alpha\sin\theta\,A\right].
\]
The two terms cancel, proving
\eqref{eq:two-flavor-instanton-source-ward-identity} to the declared local
order.

Now apply \(\partial_s^2\), \(\partial_p^2\), and \(\partial_s\partial_p\) to
the one-pair form of the Ward identity
\((4\partial_\theta+2p\partial_s-2s\partial_p)\mathcal W=0\), and then set all
sources to zero.  The three resulting identities are
\[
  4\partial_\theta\chi_s-4m_{sp}=0,\qquad
  4\partial_\theta\chi_p+4m_{sp}=0,\qquad
  4\partial_\theta m_{sp}+2\chi_s-2\chi_p=0,
\]
which are exactly
\eqref{eq:two-flavor-instanton-curvature-ward-relations}.
\end{proof}

\begin{controlledapproximation}[Tested chiral susceptibilities from the
local instanton curvature]
\label{ca:instanton-tested-chiral-susceptibility-contact}
Keep the same two-flavor source convention, but now let the scalar and
pseudoscalar sources depend on Euclidean position in a finite regulator.  Write
\[
  O_\sigma=\bar q q,\qquad
  O_{\delta_a}=\bar q\tau_a q,\qquad
  O_\eta=\bar q\ii\gamma_5q,\qquad
  O_{\pi_a}=\bar q\ii\gamma_5\tau_aq ,
\]
with the same normalization as the source coordinates
\((s_0,s_a,p_0,p_a)\) above.  For a symmetric test kernel \(F(x,y)\), define
the per-volume tested Euclidean susceptibility coordinate
\begin{equation}
  \chi_A[F]
  =
  \frac{1}{V_4}
  \left.
  \int\dd^4x\,\dd^4y\,
  F(x,y)
  \frac{\delta^2\log Z}{\delta J_A(x)\delta J_A(y)}
  \right|_{J=0},
  \qquad
  F_\Delta=
  \frac{1}{V_4}\int\dd^4x\,F(x,x).
  \label{eq:instanton-tested-susceptibility-definition}
\end{equation}
On a lattice or other cutoff regulator the integrals are the corresponding
finite sums and \(F_\Delta\) is the average diagonal value of the test kernel.
If \(F(x,y)=f(x-y)\) is translation-invariant, then \(F_\Delta=f(0)\); if the
observable is point-split so that the diagonal is removed, \(F_\Delta=0\).

The local part of the one-instanton source functional is
\[
  \log Z_{\rm inst,loc}^{(2)}
  =
  \int\dd^4x\,
  \mathcal W_{\rm inst}^{(2)}(J(x)) ,
\]
up to the derivative-expansion and finite-window remainders already declared
in ControlledApproximation~\ref{ca:two-flavor-instanton-source-susceptibility}.
Two functional derivatives of this local term produce only the regulator
diagonal:
\[
  \left.
  \frac{\delta^2\log Z_{\rm inst,loc}^{(2)}}
       {\delta J_A(x)\delta J_B(y)}
  \right|_{J=0}
  =
  C_{AB}^{\rm inst}\,\delta_\Lambda(x,y).
\]
The nonzero diagonal entries are
\[
  C_{\sigma\sigma}^{\rm inst}
  =
  C_{\pi_a\pi_a}^{\rm inst}
  =
  \kappa_{\rm inst}^{(2)}\cos\theta,
  \qquad
  C_{\delta_a\delta_a}^{\rm inst}
  =
  C_{\eta\eta}^{\rm inst}
  =
  -\kappa_{\rm inst}^{(2)}\cos\theta,
\]
and the CP-odd contact mixings are
\[
  C_{\sigma\eta}^{\rm inst}
  =
  -\kappa_{\rm inst}^{(2)}\sin\theta,
  \qquad
  C_{\delta_a\pi_b}^{\rm inst}
  =
  \kappa_{\rm inst}^{(2)}\sin\theta\,\delta_{ab}.
\]
Consequently the local instanton contribution to a tested susceptibility is
\begin{equation}
  \chi_{A,{\rm inst,loc}}[F]
  =
  C_{AA}^{\rm inst}\,F_\Delta ,
  \label{eq:instanton-tested-susceptibility-local-contact}
\end{equation}
and, at \(\theta=0\),
\begin{equation}
  \left(\chi_{\pi_a}[F]-\chi_{\delta_a}[F]\right)_{\rm inst,loc}
  =
  \left(\chi_{\sigma}[F]-\chi_{\eta}[F]\right)_{\rm inst,loc}
  =
  2\kappa_{\rm inst}^{(2)}F_\Delta .
  \label{eq:instanton-tested-u1a-contact-splitting}
\end{equation}

Thus the local 't Hooft calculation contributes the diagonal contact
coordinate of the tested correlator, not the full off-diagonal two-point
function.  A full color-singlet susceptibility difference has the
finite-regulator decomposition
\begin{equation}
\begin{aligned}
  \chi_{\pi_a-\delta_a}[F]
  &=
  2\kappa_{\rm inst}^{(2)}\cos\theta\,F_\Delta
  +
  \left(\mathsf S_{\pi_a}[F]-\mathsf S_{\delta_a}[F]\right)  \\
  &\qquad
  +
  \left(\mathsf C_{\pi_a}^{\rm ct}[F]-\mathsf C_{\delta_a}^{\rm ct}[F]\right)
  +
  R_{\rm loc}[F]+R_{\rm reg}[F]+R_{\rm IR}[F],
\end{aligned}
  \label{eq:instanton-tested-susceptibility-full-decomposition}
\end{equation}
with the analogous \(\sigma-\eta\) formula.  Here
\(\mathsf S_A[F]\) is the connected propagated, or spectral, part of the
renormalized color-singlet two-point function after the contact convention has
been chosen; \(\mathsf C_A^{\rm ct}\) records finite local counterterm
conventions; \(R_{\rm loc}\) is the omitted derivative expansion of the
instanton source kernel; \(R_{\rm reg}\) is the cutoff and volume residual;
and \(R_{\rm IR}\) contains long-distance phase information such as Goldstone
poles or screening.  In a finite box the propagated term may be extracted from
the usual Hilbert-space spectral representation of the color-singlet
operators.  The instanton calculation fixes the local \(U(1)_A\)-violating
contact coefficient in that representation.  It does not by itself determine
the spectral weights, pole residues, screening masses, or the continuum
contact convention.
\end{controlledapproximation}

\begin{proposition}[Mixed mass-source expansion of the instanton determinant]
\label{prop:instanton-mass-source-minor-expansion}
In the same \(Q=1\) QCD zero-mode sector, let \(M\) be the mass-overlap
matrix and let \(B\) be the matrix supplied by external bilinear sources or
by a specified zero-mode-paired operator insertion.  Ordinary external fermion
wave packets are represented instead by the differentiated linear-source
construction below.  With
\([N_f]=\{1,\ldots,N_f\}\), write \(M_{I^c,J^c}\) for the submatrix obtained
by deleting the rows \(I\subset [N_f]\) and columns \(J\subset [N_f]\), and
write \(B_{I,J}\) for the corresponding \(I\times J\) source submatrix, both
with increasing orders; empty determinants are \(1\).  Then
\begin{equation}
  \det(M+B)
  =
  \sum_{r=0}^{N_f}
  \sum_{\substack{I,J\subset [N_f]\\ |I|=|J|=r}}
  (-1)^{\sum_{i\in I}i+\sum_{j\in J}j}\,
  \det M_{I^c,J^c}\,
  \det B_{I,J}.
  \label{eq:instanton-mass-source-minor-expansion}
\end{equation}
Thus a correlator with \(r\) zero-mode-paired external insertions is not
obtained by appending those insertions to an already-formed vacuum
determinant.  It is the degree-\(r\) part of the same finite Berezin
determinant; the unobserved \(N_f-r\) zero-mode pairs must be saturated by the
complementary mass minor.  Equivalently, for increasing
\(I=(i_1<\cdots<i_r)\), \(J=(j_1<\cdots<j_r)\), and
\(\pi\in S_r\), the coefficient of
\(\prod_{\ell=1}^r B_{i_\ell,j_{\pi(\ell)}}\) in \(\det(M+B)\) is
\begin{equation}
  (-1)^{\sum I+\sum J}\,
  \operatorname{sgn}(\pi)\,
  \det M_{I^c,J^c}.
  \label{eq:instanton-source-differentiation-minor}
\end{equation}
For \(N_f=2\),
\begin{equation}
\begin{aligned}
  \det(M+B)
  &=
  \det M
  +M_{22}B_{11}-M_{21}B_{12}
  -M_{12}B_{21}+M_{11}B_{22}  \\
  &\hspace{2.5em}
  +B_{11}B_{22}-B_{12}B_{21}.
\end{aligned}
  \label{eq:instanton-two-flavor-mass-source-expansion}
\end{equation}
The last line is the massless four-fermion 't Hooft vertex.  The middle line
is the mixed amplitude in which one flavor is supplied by an external
zero-mode pair and the other by a mass insertion.  If a flavor is neither
externally inserted nor present in the complementary mass minor, the
instanton contribution to that correlator vanishes.
\end{proposition}

\begin{proof}
Apply the determinant expansion to \(M+B\).  Each contributing permutation
\(\sigma\in S_{N_f}\) chooses, for every row, either an \(M\)-entry or a
\(B\)-entry.  Let \(I\) be the set of rows where a \(B\)-entry was chosen and
let \(J=\sigma(I)\).  Reordering the rows and columns so that the \(B\)-block
appears in the \(I,J\) positions gives the cofactor sign
\((-1)^{\sum I+\sum J}\).  Summing over the restriction of \(\sigma\) to
\(I\) gives \(\det B_{I,J}\), while summing over the complementary
restriction gives \(\det M_{I^c,J^c}\).  This proves
\eqref{eq:instanton-mass-source-minor-expansion}.  Extracting a particular
source monomial from \(\det B_{I,J}\) gives the \(\operatorname{sgn}(\pi)\)
factor in \eqref{eq:instanton-source-differentiation-minor}.  The displayed
two-flavor formula is the specialization to \(N_f=2\).
\end{proof}

\begin{proposition}[Fermion determinant factor versus zero-mode determinant]
\label{prop:instanton-fermion-determinant-zero-mode-factorization}
Fix the same finite-regulator \(Q=1\) chart \(z=(a,\rho,U)\) for vectorlike
\(SU(N_c)\) QCD, and add a mass matrix and an even bilinear source.  Let
\(\mathcal K_{\Lambda,z}(M,B)\) be the finite Dirac bilinear between the
right and left fermion coefficient spaces in this background.  Decompose
these finite spaces into instanton zero-mode subspaces and nonzero-mode
complements.  In block form,
\begin{equation}
  \mathcal K_{\Lambda,z}(M,B)
  =
  \begin{pmatrix}
    K_{00} & K_{0{\rm n}}\\
    K_{{\rm n}0} & K_{{\rm nn}}
  \end{pmatrix},
  \label{eq:instanton-fermion-block-operator}
\end{equation}
where \(K_{00}\) acts between the zero-mode coefficient spaces and
\(K_{{\rm nn}}\) acts on the nonzero-mode complements.  If \(K_{{\rm nn}}\)
is invertible, then
\begin{equation}
  \det \mathcal K_{\Lambda,z}(M,B)
  =
  \det K_{{\rm nn}}\,
  \det\!\left(
    K_{00}-K_{0{\rm n}}K_{{\rm nn}}^{-1}K_{{\rm n}0}
  \right).
  \label{eq:instanton-fermion-schur-factorization}
\end{equation}
Thus the finite nonzero-mode fermion determinant and the zero-mode
saturation determinant are separate factors before the collective-coordinate
integral is performed.

In the leading local semiclassical approximation, where the mass/source
bilinear is restricted to the zero-mode subspace after the nonzero modes have
supplied their determinant ratio, the Schur complement in
\eqref{eq:instanton-fermion-schur-factorization} reduces to
\begin{equation}
  K_{00}^{\rm eff}(z)
  =
  M^{0}(z)+B^{0}(z),
  \qquad
  M^{0}_{ff'}(z)=\rho\,M_{ff'}(\mu),
  \qquad
  B^{0}_{ff'}(z)=\rho^3\bar\psi_{Rf}\psi_{Lf'}+\cdots .
  \label{eq:instanton-zero-mode-effective-matrix}
\end{equation}
The dots denote derivative corrections to the local source approximation and
the spin-color contractions fixed by the normalized zero-mode wavefunctions.
More generally, the Schur correction
\(-K_{0{\rm n}}K_{{\rm nn}}^{-1}K_{{\rm n}0}\) is a finite
nonzero-mode-propagator correction in the same regulator and scheme.
Consequently the light-fermion part of a one-instanton chart has the
structure
\begin{equation}
  \mathcal R_{{\rm f},\Lambda}^{\mathcal S}(z;\mu)\,
  \det K_{00}^{\rm eff}(z),
  \label{eq:instanton-fermion-nonzero-zero-factor}
\end{equation}
after trivial-sector normalization and local counterterms in the scheme
\(\mathcal S\).  The finite factor
\(\mathcal R_{{\rm f},\Lambda}^{\mathcal S}\) is the fermion nonzero-mode
determinant datum.  It is not determined by the Berezin determinant
\(\det(M^{0}+B^{0})\).  In particular, in the diagonal massive vacuum
specialization the factor \(\prod_f(m_f\rho)\) is the zero-mode saturation
determinant; the remaining finite spectral determinant is an additional
input to the physical amplitude.
\end{proposition}

\begin{proof}
At fixed cutoff the fermion path integral is a finite Berezin Gaussian, hence
equals the determinant of its finite bilinear matrix after the chosen
trivial-sector normalization.  Decompose the finite coefficient spaces into
the zero-mode subspace and its complement.  Gaussian integration over the
nonzero-mode variables first gives the Schur complement of the lower-right
block, which is exactly
\eqref{eq:instanton-fermion-schur-factorization}.  The determinant of
\(K_{{\rm nn}}\) is the fermionic part of the nonzero-mode determinant factor
in \(\mathcal W_{\Lambda}^{\rm nz}\).  The determinant of the Schur
complement is the finite Berezin integral over the zero-mode coefficients.
When the source and mass insertions are projected to the zero-mode subspace,
this second determinant is the determinant of \(M^0+B^0\), whose minor
expansion is Proposition~\ref{prop:instanton-mass-source-minor-expansion}.
If the source or mass also couples zero and nonzero modes, the same finite
Gaussian integral produces the displayed Schur correction.  Since
\(\det K_{{\rm nn}}\) multiplies every zero-mode minor, no zero-mode algebra
can determine its finite value.
\end{proof}

\begin{proposition}[Finite source-frame covariance of the light-fermion determinant]
\label{prop:instanton-light-fermion-source-frame-covariance}
In the setting of
Proposition~\ref{prop:instanton-fermion-determinant-zero-mode-factorization},
write the finite zero-mode source functional as
\begin{equation}
  \mathcal V_{\rm inst}^{0,\mathcal S}(z)
  =
  \mathcal R_{{\rm f},\Lambda}^{\mathcal S}(z;\mu)\,
  \det K_{00}^{\rm eff}(z).
  \label{eq:instanton-light-fermion-source-frame-functional}
\end{equation}
Let \(Z_R,Z_L\in {\rm GL}(N_f,\mathbb C)\) be a finite change of the right and left
zero-mode source frames, so that the same source insertion is represented by
\begin{equation}
  K_{00}^{\rm eff\,\prime}=Z_RK_{00}^{\rm eff}Z_L .
  \label{eq:instanton-zero-mode-source-frame-change}
\end{equation}
Then the same finite-regulator amplitude coordinate is represented only if
\begin{equation}
  \mathcal R_{{\rm f},\Lambda}^{\mathcal S\,\prime}
  =
  \frac{\mathcal R_{{\rm f},\Lambda}^{\mathcal S}}
       {\det Z_R\,\det Z_L},
  \qquad
  \mathcal R_{{\rm f},\Lambda}^{\mathcal S\,\prime}
  \det K_{00}^{\rm eff\,\prime}
  =
  \mathcal R_{{\rm f},\Lambda}^{\mathcal S}
  \det K_{00}^{\rm eff}.
  \label{eq:instanton-light-fermion-source-frame-covariance}
\end{equation}
If the finite scheme is also changed by a local fermion determinant counterterm
\(\Delta C_{\rm f}(z)\), then the determinant coordinate is instead multiplied
by \(\exp[-\Delta C_{\rm f}(z)]\):
\begin{equation}
  \mathcal R_{{\rm f},\Lambda}^{\mathcal S'}
  =
  \ee^{-\Delta C_{\rm f}(z)}
  \frac{\mathcal R_{{\rm f},\Lambda}^{\mathcal S}}
       {\det Z_R\,\det Z_L}.
  \label{eq:instanton-light-fermion-counterterm-frame-change}
\end{equation}
Thus a numerical value for
\(\mathcal R_{{\rm f},\Lambda}^{\mathcal S}\) is not a standalone output of
the zero-mode determinant.  It has meaning only together with the finite source
frame, local counterterm convention, and operator normalization used to extract
the physical correlator.
\end{proposition}

\begin{proof}
Equation~\eqref{eq:instanton-zero-mode-source-frame-change} and determinant
multiplicativity give
\[
  \det K_{00}^{\rm eff\,\prime}
  =
  \det Z_R\,\det K_{00}^{\rm eff}\,\det Z_L .
\]
The inverse transformation of
\(\mathcal R_{{\rm f},\Lambda}^{\mathcal S}\) in
\eqref{eq:instanton-light-fermion-source-frame-covariance} is therefore exactly
what preserves the finite product
\eqref{eq:instanton-light-fermion-source-frame-functional}.  A finite local
counterterm changes the regulated determinant density by the exponential of
minus the counterterm difference between the instanton and trivial sectors, as
in \eqref{eq:finite-regulator-instanton-measure-density}; this gives
\eqref{eq:instanton-light-fermion-counterterm-frame-change}.  The final
statement follows because source-frame changes and local counterterms act on
the finite spectral determinant factor, while the Berezin determinant records
only the zero-mode saturation algebra.
\end{proof}

\begin{controlledapproximation}[Reference-amplitude calibration of the
light-fermion determinant]
\label{ca:instanton-light-fermion-reference-calibration}
The finite light-fermion determinant factor
\(\mathcal R_{{\rm f},\Lambda}^{\mathcal S}(z;\mu)\) in
Proposition~\ref{prop:instanton-light-fermion-source-frame-covariance} can be
computed directly from the nonzero-mode spectrum in the chosen regulator, or it
can be fixed by one reference amplitude and then used as a calibrated
source-family coordinate.  This second route is predictive when the reference
projection has a declared positive noncancellation margin.

Fix one finite source frame and a collective-coordinate window
\(\mathcal U_{\rm win}\).  Let \(H_\alpha(z)\) denote the rest of the
one-instanton integrand for a gauge-invariant source channel
\(\mathcal J_\alpha\): the classical action, collective density, pure-gauge and
heavy-field determinants, zero-mode source coefficient, orientation projection,
and any declared source-to-state overlap.  Choose a reference shape
\(\overline R_{\rm f}(z)\) for the light-fermion nonzero-mode factor and write
the retained coefficient integrals
\begin{equation}
  B_\alpha
  =
  \int_{\mathcal U_{\rm win}}
    \overline R_{\rm f}(z)H_\alpha(z)\,\dd z,
  \qquad
  M_\alpha
  =
  \int_{\mathcal U_{\rm win}}
    |\overline R_{\rm f}(z)H_\alpha(z)|\,\dd z .
  \label{eq:instanton-reference-determinant-basis-integrals}
\end{equation}
If the exact one-instanton contribution in this source family has the form
\begin{equation}
  A_\alpha^{Q=1}
  =
  r_{\rm f}^{\mathcal S}B_\alpha+R_\alpha,
  \qquad
  |R_\alpha|\le E_\alpha ,
  \label{eq:instanton-reference-calibrated-family}
\end{equation}
then a reference observable \(\alpha=0\) with
\begin{equation}
  A_0^{\rm phys}
  =
  r_{\rm f}^{\mathcal S}B_0+R_0,
  \qquad
  |R_0|\le E_0,\qquad B_0\ne0
  \label{eq:instanton-reference-amplitude-input}
\end{equation}
fixes the calibrated prediction
\begin{equation}
  \widehat A_\alpha^{Q=1}
  =
  A_0^{\rm phys}\frac{B_\alpha}{B_0}.
  \label{eq:instanton-reference-amplitude-prediction}
\end{equation}
The residual propagates by the exact identity
\begin{equation}
  A_\alpha^{Q=1}-\widehat A_\alpha^{Q=1}
  =
  R_\alpha-\frac{B_\alpha}{B_0}R_0,
  \qquad
  \left|A_\alpha^{Q=1}-\widehat A_\alpha^{Q=1}\right|
  \le
  E_\alpha+\left|\frac{B_\alpha}{B_0}\right|E_0 .
  \label{eq:instanton-reference-calibration-error}
\end{equation}
Thus the reference channel must supply a noncancellation margin
\begin{equation}
  \kappa_0=\frac{|B_0|}{M_0}>0,
  \qquad
  \left|\frac{B_\alpha}{B_0}\right|
  \le
  \frac{M_\alpha}{\kappa_0 M_0}.
  \label{eq:instanton-reference-noncancellation-margin}
\end{equation}

This is the finite-amplitude meaning of a light-fermion determinant
normalization.  A finite source-frame change multiplies every \(B_\alpha\) in
the same family by the corresponding zero-mode determinant factor and changes
\(r_{\rm f}^{\mathcal S}\) inversely, so the calibrated ratio in
\eqref{eq:instanton-reference-amplitude-prediction} is invariant for reference
and target channels kept in the same frame and local counterterm
scheme.  The controlled scope of the calibration is one overall determinant
coordinate; residual \(z\)-dependence of the determinant, the omitted size
window, the Euclidean-to-physical bridge, and the long-distance spectral matrix
element require separate estimates.  If the reference \(B_0\) is small because
of source or orientation cancellations, the uncertainty is amplified by
\eqref{eq:instanton-reference-calibration-error}; precision use for the
remaining channels therefore requires the margin in
\eqref{eq:instanton-reference-noncancellation-margin}.
\end{controlledapproximation}

\begin{controlledapproximation}[Finite-scheme transport of an instanton
determinant constant]
\label{ca:instanton-determinant-scheme-transport}
A quoted instanton determinant constant is meaningful only as part of a
finite scheme datum.  Let \(\mathcal S\) and \(\mathcal S'\) be two
mass-independent finite schemes for the same \(Q=1\) amplitude channel, with
the same classical chart and the same physical external observable.  Write
\[
  X_{\mathcal S}(\mu)=\frac{8\pi^2}{g_{\mathcal S}^2(\mu)} .
\]
Suppose the finite coupling-coordinate change, source-zero-mode coordinate
change, and final operator/source projection are
\begin{equation}
  X_{\mathcal S'}=X_{\mathcal S}+\delta_X+r_X,\qquad
  B_{\mathcal S'}=z_B\,\ee^{r_B}B_{\mathcal S},\qquad
  O_{\mathcal S'}=z_O\,\ee^{r_O}O_{\mathcal S},
  \label{eq:instanton-finite-scheme-map}
\end{equation}
with \(z_B,z_O\ne0\).  Here \(B_{\mathcal S}\) denotes the retained
zero-mode/source integral or finite-window source tensor, while
\(O_{\mathcal S}\) is the matrix element, OPE-basis element, or physical
projection coordinate multiplying the instanton coefficient.  Logarithms are
taken in the chosen finite coordinate patch; for a positive activity bound one
uses the same formulas with logarithms of absolute values.  For a purely
Euclidean source coefficient one sets \(z_O=1\).  For a mass-saturated vacuum
term, \(z_B\) is the determinant of the finite mass-coordinate change; for a
four-source amplitude it is the product of the right and left source-frame
determinants.

If the leading one-instanton contribution is represented in scheme
\(\mathcal S\) by
\[
  \mathcal A_{\mathcal S}^{\rm lead}
  =
  C_{\mathcal S}^{\rm det}
  \ee^{-X_{\mathcal S}}
  B_{\mathcal S}O_{\mathcal S},
\]
then the transported determinant constant in \(\mathcal S'\) is
\begin{equation}
  C_{\mathcal S'}^{\rm det,lead}
  =
  \frac{\ee^{\delta_X}}{z_Bz_O}\,
  C_{\mathcal S}^{\rm det}.
  \label{eq:instanton-determinant-finite-scheme-transport}
\end{equation}
Indeed, inserting
\eqref{eq:instanton-determinant-finite-scheme-transport} and dropping the
remainders \(r_X,r_B,r_O\) gives the same leading amplitude coordinate.
Thus one may not change from a Pauli--Villars coupling to another finite
coupling convention, or rescale the external source/operator frame, while
reusing the old numerical determinant constant.

More generally, suppose the actually supplied constant satisfies the
logarithmic transport bound
\begin{equation}
  \left|
  \log C_{\mathcal S'}^{\rm det}
  -
  \log C_{\mathcal S}^{\rm det}
  -
  \delta_X+\log z_B+\log z_O
  \right|
  \le\epsilon_C,
  \label{eq:instanton-determinant-scheme-log-bound}
\end{equation}
and that
\[
  |r_X|\le\epsilon_X,\qquad |r_B|\le\epsilon_B,\qquad
  |r_O|\le\epsilon_O .
\]
Then the leading amplitude coordinates in the two schemes obey
\begin{equation}
  \left|
  \frac{\mathcal A_{\mathcal S'}^{\rm lead}}
       {\mathcal A_{\mathcal S}^{\rm lead}}
  -1
  \right|
  \le
  \exp(\epsilon_C+\epsilon_X+\epsilon_B+\epsilon_O)-1 .
  \label{eq:instanton-determinant-scheme-amplitude-bound}
\end{equation}
This bound is a scheme-transport residual.  It is distinct from the
proper-time determinant residual, the rank/source residual, and the endpoint
tail.  A finite determinant constant in a declared scheme is therefore not a
decorative number attached to the BPST moduli measure; it is an amplitude
coordinate whose compensation against the coupling, source, and operator
coordinates must be checked.
\end{controlledapproximation}

\begin{proposition}[Mass/source RG transport of the instanton vertex]
\label{prop:instanton-mass-source-rg-transport}
Work in a mass-independent renormalization scheme in the same \(Q=1\)
vectorlike QCD chart.  Let \(\gamma_m(g)\) be the anomalous dimension of the
quark mass coordinate in the convention
\begin{equation}
  \mu\frac{\dd M}{\dd\mu}=-\gamma_m(g)M,
  \qquad
  \mu\frac{\dd J}{\dd\mu}=-\gamma_m(g)J,
  \label{eq:instanton-mass-source-running}
\end{equation}
where \(J\) is a renormalized source coordinate coupled to the same chiral
scalar density as the mass matrix.  Let \(M^0(z)=\rho M\), and let
\(J^0(z)\) be the zero-mode projection of \(J\), a linear function of \(J\)
with the same anomalous transport as \(J\).  Let
\(\mathcal R_{\rm f}^{\mathcal S}(z;\mu)\) denote the finite light-fermion
nonzero-mode determinant factor in the source convention of
Proposition~\ref{prop:instanton-fermion-determinant-zero-mode-factorization}.
If, at the logarithmic accuracy under consideration,
\begin{equation}
  \left(\mu\partial_\mu+\beta(g)\partial_g\right)
  \log \mathcal R_{\rm f}^{\mathcal S}(z;\mu)
  =
  N_f\,\gamma_m(g),
  \label{eq:instanton-fermion-factor-mass-anomalous-rge}
\end{equation}
then the zero-mode source functional
\begin{equation}
  \mathcal V_{\rm inst}^{0}(z;M,J,\mu)
  =
  \mathcal R_{\rm f}^{\mathcal S}(z;\mu)\,
  \det\!\left(M^0(z)+J^0(z)\right)
  \label{eq:instanton-renormalized-source-functional}
\end{equation}
obeys
\begin{equation}
  \left[
    \mu\partial_\mu+\beta(g)\partial_g
    -\gamma_m(g)
    \sum_{f,f'}
    \left(
      M_{ff'}\frac{\partial}{\partial M_{ff'}}
      +J_{ff'}\frac{\partial}{\partial J_{ff'}}
    \right)
  \right]
  \mathcal V_{\rm inst}^{0}
  =
  0 .
  \label{eq:instanton-source-functional-rge}
\end{equation}
Thus the mass/source determinant is not an RG-invariant number by itself.
The source functional is covariant only after the finite nonzero-mode
determinant factor and the source renormalization convention are included.
The symbol \(J\) denotes the source coordinate before differentiation.  The
scalar-density operator obtained after differentiating the generating
functional uses the corresponding operator normalization; in the local
operator convention that differentiated insertion has the
\(B^0_{ff'}=\rho^3\bar\psi_{Rf}\psi_{Lf'}+\cdots\) normalization of
\eqref{eq:instanton-zero-mode-effective-matrix}.

In particular, for diagonal masses and \(J=0\), the anomalous-dimension part
of the vacuum factor
\[
  \mathcal R_{\rm f}^{\mathcal S}(\mu)\prod_{f=1}^{N_f}m_f(\mu)
\]
is scale invariant at the same logarithmic accuracy.  The phase
\(\theta+\arg\det M\) is unaffected by the positive real mass
renormalization factor; it is fixed instead by the anomalous chiral rotation
described in \eqref{eq:qcd-physical-theta}.  After differentiating
\eqref{eq:instanton-renormalized-source-functional} \(r\) times with respect
to \(J\) and setting \(J=0\), the resulting mixed mass/source minor has the
RG covariance of an \(r\)-fold insertion of the renormalized scalar density,
rather than the covariance of a vacuum activity.
\end{proposition}

\begin{proof}
The matrix \(M^0(z)+J^0(z)\) is homogeneous of degree one in the source
coordinates \(M\) and \(J\), while the projection factors depending on
\(\rho\), \(U\), and the normalized zero-mode wavefunctions are collective
coordinate data and are not acted on by the source anomalous dimension.  The
determinant is therefore homogeneous of degree \(N_f\):
\[
  \sum_{f,f'}
  \left(
    M_{ff'}\frac{\partial}{\partial M_{ff'}}
    +J_{ff'}\frac{\partial}{\partial J_{ff'}}
  \right)
  \det(M^0+J^0)
  =
  N_f\det(M^0+J^0).
\]
Combining this identity with
\eqref{eq:instanton-fermion-factor-mass-anomalous-rge} gives
\eqref{eq:instanton-source-functional-rge}.  Setting \(J=0\) gives the
vacuum statement.  Differentiating the source-functional equation \(r\) times
with respect to \(J\) transfers \(r\) source anomalous-dimension terms from
the transport operator to the differentiated coefficient, which is the usual
Callan--Symanzik covariance of \(r\) renormalized scalar-density insertions.
The phase statement follows from a separate observation.  In a vectorlike
mass-independent scheme, multiplicative mass renormalization changes the
magnitude of \(M\).  The anomalous flavor-singlet chiral rotation changes
\(\arg\det M\) and \(\theta\) in opposite directions.
\end{proof}

\begin{proposition}[Proper-time determinant and zero-mode source functional]
\label{prop:instanton-proper-time-four-fermion-amplitude}
Fix a finite-regulator \(Q=1\) chart \(z=(a,\rho,U)\), a background gauge
slice, and a proper-time determinant prescription with renormalization scale
\(\mu\).  For each positive nonzero-mode operator \(P_{\Lambda,z}\) in the
instanton background, compared with the corresponding operator
\(P_{\Lambda,0}\) in the trivial sector, define the regulated logarithmic
determinant ratio by
\begin{equation}
  K_\Lambda(P_z,P_0;\mu)
  =
  -\int_{\epsilon_\Lambda}^{\infty}
    \frac{\dd t}{t}\,r(t\mu^2)
    \left(
      \operatorname{Tr}' e^{-tP_{\Lambda,z}}
      -
      \operatorname{Tr} e^{-tP_{\Lambda,0}}
    \right)
  +C_{\Lambda,P}^{\rm loc}(z;\mu),
  \label{eq:proper-time-instanton-determinant-ratio}
\end{equation}
where \(r\) is a smooth cutoff profile with \(r(0)=1\), the prime removes
collective-coordinate, residual-gauge, and fermion zero modes as appropriate,
and \(C_{\Lambda,P}^{\rm loc}\) is the chosen local subtraction in the
instanton background relative to the trivial background.  The determinant
density in \eqref{eq:finite-regulator-instanton-measure-density} is then
\begin{equation}
  \log \mathcal W_{\Lambda,{\rm pt}}^{\rm nz}(z;\mu)
  =
  K_\Lambda(M_z^{\rm gh},M_0^{\rm gh};\mu)
  -\frac12 K_\Lambda(L_z^{\rm bos},L_0^{\rm bos};\mu)
  +\sum_R K_\Lambda^{\rm pf,R}(z,0;\mu),
  \label{eq:proper-time-instanton-Wnz}
\end{equation}
with \(K_\Lambda^{\rm pf,R}\) the Pfaffian or determinant-line version of the
same nonzero-mode construction for the fermion multiplet \(R\).

For two massless Dirac flavors there are two distinct source constructions.
An even bilinear source supplies an independent finite matrix \(B(z)\) in the
zero-mode action \(\sum_{f,f'}\widetilde\lambda_f B_{ff'}(z)\lambda_{f'}\);
its zero-mode factor is \(\det B(z)\), as in
Proposition~\ref{prop:qcd-thooft-determinant-zero-mode-amplitude}.  Ordinary
external fermion wave packets are instead introduced by odd generating
coordinates.  Let \(\bar\chi_A,\chi_B\), \(A,B=1,2\), be the four external
source coordinates and set
\begin{equation}
  \alpha_f(z)=\sum_{A=1}^{2}\bar\chi_A R_{Af}(z),
  \qquad
  \beta_f(z)=\sum_{B=1}^{2}L_{fB}(z)\chi_B,
  \label{eq:zero-mode-linear-source-overlaps}
\end{equation}
where \(R_{Af}\) and \(L_{fB}\) are the zero-mode overlaps of the \(A\)th
right-handed and \(B\)th left-handed external slots with the normalized
instanton zero-mode wavefunction.  With the same Berezin orientation used in
\eqref{eq:thooft-vertex-berezin-determinant}, the differentiated source
coefficient is
\begin{equation}
  [\bar\chi_1\bar\chi_2\chi_1\chi_2]\,
  \prod_{f=1}^{2}\alpha_f(z)\prod_{f=1}^{2}\beta_f(z)
  =
  \det R(z)\,\det L(z).
  \label{eq:zero-mode-differentiated-four-slot-coefficient}
\end{equation}
Thus the one-instanton generating-functional contribution is
\begin{equation}
\begin{aligned}
  \mathcal G_{\eta,\Lambda,Q=1}^{(4)}[\bar\chi,\chi]
  &=
  \int_{\mathcal U_{Q=1}}
  \exp[-S_\Lambda(A_z)+\ii\theta]\,
  \mathcal W_{\Lambda,{\rm pt}}^{\rm nz}(z;\mu)\,
  \prod_{f=1}^{2}\alpha_f(z)\prod_{f=1}^{2}\beta_f(z)\,
  \dd\nu_{\Lambda,\mathcal M}(z),\\
  G^{(4)}_{\eta,\Lambda,Q=1}
  &=
  [\bar\chi_1\bar\chi_2\chi_1\chi_2]\,
  \mathcal G_{\eta,\Lambda,Q=1}^{(4)}[\bar\chi,\chi].
\end{aligned}
  \label{eq:proper-time-four-fermion-instanton-amplitude}
\end{equation}
A c-number matrix formed as a single outer product \(u_fv_{f'}\) is not this
four-slot coefficient; for \(N_f=2\) its determinant vanishes.  The
small-\(t\) expansion of
\eqref{eq:proper-time-instanton-determinant-ratio} supplies the local
counterterms and the universal \((\mu\rho)^{b_0}\) factor, while the finite
remainder fixes the scheme-dependent numerical constant in the selected
source functional.  Thus the physics calculation consists of the regulated
fluctuation determinant, the zero-mode source differentiation, and the
collective-coordinate integral as a single object.
\end{proposition}

\begin{proof}
For a finite positive operator \(P\), the identity
\(\log\det P=-\int_{\epsilon}^{\infty}\dd t\,t^{-1}\operatorname{Tr}e^{-tP}\)
holds up to the subtraction convention at the lower endpoint.  Applying this
identity to the instanton and trivial-sector operators and adding the chosen
local counterterm gives
\eqref{eq:proper-time-instanton-determinant-ratio}.  Bosonic Gaussian
integration contributes the exponent \(-1/2\), ghost Berezin integration
contributes \(+1\), and fermionic nonzero modes contribute the Pfaffian or
determinant-line factor, giving
\eqref{eq:proper-time-instanton-Wnz}.  Zero modes have already been removed
from the traces and determinants, so their finite Berezin integral is exactly
the calculation in
Proposition~\ref{prop:qcd-thooft-determinant-zero-mode-amplitude} for an even
bilinear source matrix.  For linear external sources, expanding
\(\prod_f\alpha_f\prod_f\beta_f\) and extracting the
\(\bar\chi_1\bar\chi_2\chi_1\chi_2\) coefficient gives
\(\det R\,\det L\); this is the source-differentiated correlator coefficient,
not the determinant of a single c-number outer product.  Multiplication by the
collective-coordinate density and the classical instanton exponential gives
\eqref{eq:proper-time-four-fermion-instanton-amplitude}.  Finally, the
heat-kernel expansion of the trace difference at small \(t\) separates local
counterterms, the coefficient of \(\log(\mu\rho)\), and the finite
determinant constant; RG invariance identifies the universal logarithmic
coefficient with \(b_0\) in the displayed one-loop density.
\end{proof}

\begin{controlledapproximation}[Proper-time determinant residual on an amplitude window]
\label{ca:instanton-proper-time-determinant-residual-window}
The proper-time determinant in
Proposition~\ref{prop:instanton-proper-time-four-fermion-amplitude} becomes
useful for a physical source amplitude only after its finite remainder is
controlled uniformly on the same collective-coordinate window as the
zero-mode and source estimates.  Fix a window \(\mathcal U_{\rm win}\) in a
one-instanton chart and write, after the local heat-kernel subtraction and
the universal \(b_0\log(\mu\rho)\) term have been separated,
\[
  \log\mathcal W_{\Lambda,{\rm pt}}^{\rm nz}(z;\mu)
  =
  \log\mathcal W_{{\rm lead}}^{\rm nz}(z;\mu)
  +\delta_{\rm det}(z),
  \qquad z\in\mathcal U_{\rm win}.
\]
For each fluctuation block \(j\) let \(c_j\) be its logarithmic weight:
\[
  c_{\rm gh}=1,\qquad c_{\rm bos}=-\frac12,\qquad
  c_{\rm pf,R}=1
\]
with the Pfaffian or determinant-line pairing convention already fixed.  A
determinant-residual bound on \(\mathcal U_{\rm win}\) is a collection
of trace-remainder bounds
\[
  \left|
    \int_{\epsilon_\Lambda}^{\infty}
    \frac{\dd t}{t}\,r(t\mu^2)\,
    \delta T_{j,\Lambda}(t,z)
  \right|
  \le \epsilon_j,
  \qquad z\in\mathcal U_{\rm win},
\]
where \(\delta T_{j,\Lambda}\) is the remaining trace difference after the
declared local subtraction and leading determinant datum have been removed.
It gives the uniform logarithmic bound
\begin{equation}
  |\delta_{\rm det}(z)|
  \le
  \epsilon_{\rm det}:=\sum_j |c_j|\epsilon_j .
  \label{eq:instanton-proper-time-determinant-log-bound}
\end{equation}
Consequently the multiplicative determinant error satisfies
\begin{equation}
  \left|
    \mathcal W_{\Lambda,{\rm pt}}^{\rm nz}(z;\mu)
    -
    \mathcal W_{{\rm lead}}^{\rm nz}(z;\mu)
  \right|
  \le
  \left|\mathcal W_{{\rm lead}}^{\rm nz}(z;\mu)\right|
  \left(e^{\epsilon_{\rm det}}-1\right).
  \label{eq:instanton-proper-time-determinant-multiplicative-bound}
\end{equation}
Thus the determinant part of the amplitude error budget obeys
\begin{equation}
  |R_{\rm det}|
  \le
  \left(e^{\epsilon_{\rm det}}-1\right)
  \int_{\mathcal U_{\rm win}}
  |\mathcal K_{\rm lead}(z)|\,\dd z ,
  \label{eq:instanton-proper-time-determinant-amplitude-bound}
\end{equation}
provided the leading kernel contains the same zero-mode source coefficient,
collective density, and local counterterm convention.

This is the finite-regulator form of the fluctuation calculation hidden in a
``determinant constant.''  A signed cancellation among the ghost, bosonic,
and fermionic trace remainders is weaker than
\eqref{eq:instanton-proper-time-determinant-log-bound}; the amplitude bound
uses absolute logarithmic control because the determinant multiplies the
zero-mode source coefficient pointwise over the size and orientation window.
The finite constant may be pulled outside the instanton size integral only
when the same residual bound also controls the residual \(z\)-dependence of the
finite determinant over that window.
\end{controlledapproximation}

\begin{proposition}[Instanton zero-mode density form factor and local limit]
\label{prop:instanton-zero-mode-tail-local-limit}
In a BPST \(Q=1\) background, after the spin and color tensors of one
fundamental zero-mode line are factored out, the radial scalar envelope which
weights a local mass, Yukawa, or bilinear source is
\begin{equation}
  h_\rho(y)
  =
  \frac{2\rho^2}{\pi^2(y^2+\rho^2)^3},
  \qquad
  \int_{\mathbb R^4}h_\rho(y)\,\dd^4y=1 .
  \label{eq:instanton-zero-mode-envelope}
\end{equation}
Thus a zero-mode-paired source matrix entry has the form
\begin{equation}
  B_{ff'}^{(s)}(a,\rho,U)
  =
  \rho^{d_s}
  \int_{\mathbb R^4}
  h_\rho(y)\,
  \mathcal B_{ff'}(a+y;U)\,\dd^4y ,
  \label{eq:instanton-zero-mode-source-convolution}
\end{equation}
where \(d_s=1\) for a mass source and \(d_s=3\) for the local quark-bilinear
source in \eqref{eq:qcd-thooft-vertex-leading-local}.  Equivalently, with
the Fourier convention
\[
  \widehat h_\rho(q)=
  \int_{\mathbb R^4}\ee^{\ii q\cdot y}h_\rho(y)\,\dd^4y ,
\]
one has the exact form factor
\begin{equation}
  \widehat h_\rho(q)
  =
  zK_1(z),
  \qquad z=\rho |q|,
  \label{eq:instanton-zero-mode-momentum-form-factor}
\end{equation}
where \(K_1\) is the modified Bessel function.  Hence a zero-mode-paired
source with Fourier momentum \(q\) is not represented by the local vertex
until \(z\ll1\).
The small-momentum expansion is
\begin{equation}
  zK_1(z)
  =
  1+
  \frac{z^2}{2}
  \left[
    \log\frac{z}{2}+\gamma_{\rm E}-\frac12
  \right]
  +O(z^4\log z),
  \label{eq:instanton-zero-mode-form-factor-small-momentum}
\end{equation}
so the first nonlocal correction is \(q^2\rho^2\log(|q|\rho)\), not an
ordinary analytic Taylor coefficient.  At \(z\gg1\),
\begin{equation}
  zK_1(z)
  =
  \sqrt{\frac{\pi z}{2}}\,
  \ee^{-z}
  \left(1+\frac{3}{8z}+O(z^{-2})\right),
  \label{eq:instanton-zero-mode-form-factor-large-momentum}
\end{equation}
which displays the suppression of source components varying on scales
shorter than the instanton size.  This is the density form factor for a local
bilinear or mass-source insertion.  A differentiated correlator with four
ordinary external fermion slots instead uses the Fourier transforms of the
individual zero-mode wavefunctions in those slots; their product reduces to
the local bilinear density only after the slot wavefunctions are fused into a
bilinear source and the local limit is taken.  In position space, for
\(U_R=R/\rho\),
\begin{equation}
  \int_{|y|<R}h_\rho(y)\,\dd^4y
  =
  \frac{U_R^4}{(1+U_R^2)^2},
  \qquad
  \int_{|y|>R}h_\rho(y)\,\dd^4y
  =
  \frac{1+2U_R^2}{(1+U_R^2)^2}.
  \label{eq:instanton-zero-mode-tail-mass}
\end{equation}
The first correction to a local Taylor expansion is more delicate.  The
truncated second moment is
\begin{equation}
  M_2(R)
  :=
  \int_{|y|<R}|y|^2h_\rho(y)\,\dd^4y
  =
  2\rho^2
  \left[
    \log(1+U_R^2)
    +\frac{2}{1+U_R^2}
    -\frac{1}{2(1+U_R^2)^2}
    -\frac32
  \right],
  \label{eq:instanton-zero-mode-second-moment}
\end{equation}
and rotational invariance gives
\[
  \int_{|y|<R}y^\mu y^\nu h_\rho(y)\,\dd^4y
  =
  \frac{\delta^{\mu\nu}}4M_2(R).
\]
Consequently the leading local term in
\eqref{eq:instanton-zero-mode-source-convolution} is
\(\rho^{d_s}\mathcal B_{ff'}(a;U)\), with a tail bounded by the second
quantity in \eqref{eq:instanton-zero-mode-tail-mass}.  But the would-be
two-derivative coefficient grows as
\(4\rho^2\log(R/\rho)+O(\rho^2)\) as \(R/\rho\to\infty\).  The local
't Hooft vertex is therefore the zeroth adiabatic term of the zero-mode
source convolution; higher-derivative terms require the external wave-packet,
mass, or infrared regulator that cuts off the zero-mode tail.  In momentum
space the same statement is the logarithm in
\eqref{eq:instanton-zero-mode-form-factor-small-momentum}.
\end{proposition}

\begin{proof}
The normalized radial envelope in \eqref{eq:instanton-zero-mode-envelope} is
the scalar part of the product of two normalized fundamental zero-mode
wavefunctions in the BPST background.  Using
\(\dd^4y=2\pi^2r^3\dd r\) and \(u=r/\rho\),
\[
  \int_{\mathbb R^4}h_\rho(y)\,\dd^4y
  =
  4\int_0^\infty\frac{u^3}{(1+u^2)^3}\,\dd u
  =
  1.
\]
For the Fourier transform, use the Schwinger representation
\[
  \frac1{(y^2+\rho^2)^3}
  =
  \frac12\int_0^\infty s^2\ee^{-s(y^2+\rho^2)}\,\dd s .
\]
Then
\[
\begin{aligned}
  \widehat h_\rho(q)
  &=
  \rho^2\int_0^\infty
  \exp\!\left[-\rho^2s-\frac{q^2}{4s}\right]\dd s  \\
  &=
  \rho |q|\,K_1(\rho |q|),
\end{aligned}
\]
where the last step is the standard integral representation of the modified
Bessel function.  The small-\(z\) and large-\(z\) formulas are the
corresponding standard asymptotic expansions of \(K_1\); the logarithmic
small-\(z\) term is the momentum-space reflection of the logarithmically
divergent second moment computed below.

The same integral cut off at \(U_R\) gives
\[
  4\int_0^{U_R}\frac{u^3}{(1+u^2)^3}\,\dd u
  =
  1-\frac{2}{1+U_R^2}
  +\frac{1}{(1+U_R^2)^2}
  =
  \frac{U_R^4}{(1+U_R^2)^2},
\]
and subtracting from one gives the tail formula in
\eqref{eq:instanton-zero-mode-tail-mass}.  For the second moment,
\[
  M_2(R)
  =
  4\rho^2\int_0^{U_R}\frac{u^5}{(1+u^2)^3}\,\dd u .
\]
With \(t=1+u^2\), this antiderivative is
\[
  2\rho^2
  \left[
    \log t+\frac2t-\frac{1}{2t^2}-\frac32
  \right]_{t=1}^{1+U_R^2},
\]
which is \eqref{eq:instanton-zero-mode-second-moment}.  Rotational
invariance in four Euclidean dimensions gives the displayed
\(\delta^{\mu\nu}/4\) tensor moment.  The Taylor expansion of
\(\mathcal B(a+y;U)\) then yields the zeroth local term and the
\((M_2(R)/8)\partial^2\mathcal B(a;U)\) contribution inside the ball
\(|y|<R\), while the exterior contribution is controlled by the tail mass
times the appropriate supremum norm of the source.  Since
\(\log(1+U_R^2)=2\log(R/\rho)+o(1)\), the full second moment is
logarithmically infrared sensitive.
\end{proof}

\begin{proposition}[Individual BPST zero-mode slot form factor]
\label{prop:instanton-individual-zero-mode-form-factor}
For an ordinary external fermion slot one must use the Fourier transform of a
single normalized BPST zero-mode wavefunction, not the density form factor
\(\widehat h_\rho\) of
Proposition~\ref{prop:instanton-zero-mode-tail-local-limit}.  In the standard
singular-gauge momentum-space convention for a charge-one instanton
\cite{tHooft1976Pseudoparticle,DiakonovPetrov1986InstantonVacuum}, the
spin-color tensor and the scalar radial function separate as
\begin{equation}
  \widetilde\psi_{0,i}^{\alpha}(p;\rho,U)
  =
  \sqrt2\,\phi_\rho'(|p|)
  \bigl(\widehat p\,\epsilon U\bigr)_{i}^{\alpha},
  \qquad
  \phi_\rho(p)
  =
  \pi\rho^2
  \bigl[I_0(t)K_0(t)-I_1(t)K_1(t)\bigr],
  \quad
  t=\frac{\rho |p|}{2}.
  \label{eq:instanton-momentum-zero-mode-wavefunction}
\end{equation}
Here \(I_\nu\) and \(K_\nu\) are modified Bessel functions, \(U\) is the color
orientation, and \(\widehat p=p/|p|\).  After the adjacent free Dirac operator
and fixed Fourier-normalization constants are assigned to the external-slot
amputation convention, the dimensionless scalar slot form factor is
\begin{equation}
  F_{\rm zm}(t)
  :=
  -t\frac{\dd}{\dd t}
  \bigl[I_0(t)K_0(t)-I_1(t)K_1(t)\bigr]
  =
  2t\bigl[I_0(t)K_1(t)-I_1(t)K_0(t)\bigr]
  -2I_1(t)K_1(t).
  \label{eq:instanton-individual-zero-mode-form-factor}
\end{equation}
It is normalized by
\begin{equation}
  F_{\rm zm}(0)=1,
  \label{eq:instanton-individual-zero-mode-form-factor-local}
\end{equation}
while at large momentum
\begin{equation}
  F_{\rm zm}(t)
  =
  \frac{3}{4t^3}+O(t^{-5})
  =
  \frac{6}{(\rho |p|)^3}+O((\rho |p|)^{-5}).
  \label{eq:instanton-individual-zero-mode-form-factor-large}
\end{equation}
Thus the individual external zero-mode slot has a power tail.  The exponential
large-momentum suppression in
\eqref{eq:instanton-zero-mode-form-factor-large-momentum} belongs to the fused
bilinear density \(h_\rho\), not to a single differentiated fermion slot.
\end{proposition}

\begin{proof}
The first formula is the standard Fourier-Bessel transform of the normalized
singular-gauge fundamental zero mode after its hedgehog spin-color tensor has
been extracted.  Differentiating
\(\Phi(t)=I_0(t)K_0(t)-I_1(t)K_1(t)\) and using
\[
  I_0'=I_1,\qquad
  I_1'=I_0-\frac{I_1}{t},\qquad
  K_0'=-K_1,\qquad
  K_1'=-K_0-\frac{K_1}{t}
\]
gives the second expression in
\eqref{eq:instanton-individual-zero-mode-form-factor}.  The small-\(t\)
expansions \(I_0(t)=1+O(t^2)\), \(I_1(t)=t/2+O(t^3)\), and
\(K_1(t)=t^{-1}+O(t\log t)\) give \(F_{\rm zm}(0)=1\).  The standard large-\(t\)
asymptotic expansions of \(I_\nu\) and \(K_\nu\) show that the constant,
\(t^{-1}\), and \(t^{-2}\) terms cancel in
\eqref{eq:instanton-individual-zero-mode-form-factor}; the first nonzero term
is \(3/(4t^3)\).
\end{proof}

\begin{proposition}[Amputation and source typing of the instanton kernel]
\label{prop:instanton-external-leg-amputation-kernel}
Work in the finite-regulator \(Q=1\) chart of
Proposition~\ref{prop:instanton-proper-time-four-fermion-amplitude}, and
suppose an auxiliary infrared deformation or source scheme has been chosen in
which the external quark fields used to probe the kernel have isolated
nonzero external pole-residue factors.  For differentiated linear source
slots, let \(Z^R_A(p_{R,A})\) and \(Z^L_B(p_{L,B})\) be the adjacent
unamputated pole residues for the two right-handed and two left-handed slots.
After the spin and color external covectors are fixed, write the zero-mode
overlap matrices as
\begin{equation}
\begin{aligned}
  R^{\rm unamp}_{Af}(z)
  &=
  Z^R_A(p_{R,A})\,R^{\rm amp}_{Af}(z),\\
  L^{\rm unamp}_{fB}(z)
  &=
  L^{\rm amp}_{fB}(z)\,Z^L_B(p_{L,B}),
  \qquad A,B,f=1,2 .
\end{aligned}
  \label{eq:instanton-amputated-linear-source-matrices}
\end{equation}
The four-slot coefficient from
\eqref{eq:zero-mode-differentiated-four-slot-coefficient} obeys
\begin{equation}
  \det R^{\rm unamp}(z)\det L^{\rm unamp}(z)
  =
  \left(\prod_{A=1}^{2}Z^R_A\right)
  \left(\prod_{B=1}^{2}Z^L_B\right)
  \det R^{\rm amp}(z)\det L^{\rm amp}(z).
  \label{eq:instanton-source-determinant-amputation}
\end{equation}
Consequently the source-amputated differentiated four-fermion kernel is
\begin{equation}
  \mathcal M^{(4),{\rm amp}}_{\Lambda,Q=1}
  =
  \int_{\mathcal U_{Q=1}}
  \exp[-S_\Lambda(A_z)+\ii\theta]\,
  \mathcal W_{\Lambda,{\rm pt}}^{\rm nz}(z;\mu)\,
  \det R^{\rm amp}(z)\det L^{\rm amp}(z)\,
  \dd\nu_{\Lambda,\mathcal M}(z).
  \label{eq:instanton-amputated-four-fermion-kernel}
\end{equation}
Amputation removes only the adjacent external propagator residues.  It does
not remove the individual instanton zero-mode Fourier overlaps, the spin-color
orientation tensors, the nonzero-mode determinant, or the
collective-coordinate integral.

For an independent even bilinear source matrix the source type is different.
If \(B^{\rm unamp}=D_RB^{\rm amp}D_L\), with \(D_R,D_L\) diagonal residue
matrices in flavor space, then
\begin{equation}
  \det B^{\rm unamp}
  =
  \det D_R\,\det B^{\rm amp}\,\det D_L .
  \label{eq:instanton-bilinear-source-determinant-amputation}
\end{equation}
For a local bilinear source carrying momentum \(q_{ff'}\), the BPST density
form factor of Proposition~\ref{prop:instanton-zero-mode-tail-local-limit}
gives a representative entry
\begin{equation}
  B^{\rm amp}_{ff'}(a,\rho,U)
  =
  \rho^3
  \ee^{\ii q_{ff'}\cdot a}\,
  \widehat h_\rho(q_{ff'})\,
  \mathcal T_{ff'}(U;q_{ff'}),
  \qquad
  q_{ff'}=p_{R,f}-p_{L,f'} .
  \label{eq:instanton-amputated-source-matrix}
\end{equation}
This bilinear-source formula is not the determinant of a single c-number
outer product of external wave packets.  Ordinary external fermions use
\eqref{eq:instanton-amputated-linear-source-matrices} and the source
differentiation in
\eqref{eq:zero-mode-differentiated-four-slot-coefficient}.
\end{proposition}

\begin{proof}
The linear source overlaps in \eqref{eq:zero-mode-linear-source-overlaps}
attach each external slot to one normalized zero-mode wavefunction.  The
chosen infrared scheme supplies one adjacent pole residue per external slot;
removing those residues gives
\eqref{eq:instanton-amputated-linear-source-matrices}.  Applying the
determinant to the right-slot matrix and to the left-slot matrix gives
\eqref{eq:instanton-source-determinant-amputation}.  Multiplying the
unamputated differentiated Green kernel by the inverse product of the same
four residues therefore gives
\eqref{eq:instanton-amputated-four-fermion-kernel}.

For an even bilinear source, the zero-mode action already contains a matrix
entry \(B_{ff'}\).  Row and column residue factors then multiply that matrix
on the left and right, so ordinary determinant multiplicativity gives
\eqref{eq:instanton-bilinear-source-determinant-amputation}.  The density
form factor in \eqref{eq:instanton-amputated-source-matrix} comes from the
product of two zero-mode wavefunctions in a fused bilinear source.  Linear
external slots are differentiated before the sources are set to zero, so a
single rank-one c-number matrix never appears as the four-fermion coefficient.
\end{proof}

\begin{proposition}[Plane-wave assembly of the one-instanton four-fermion
amplitude]
\label{prop:instanton-plane-wave-four-fermion-assembly}
Stay in the \(N_f=2\) source-amputated setup of
Proposition~\ref{prop:instanton-external-leg-amputation-kernel}, and suppose
the center coordinate factorizes from the regulated collective-coordinate
density:
\[
  \dd\nu_{\Lambda,\mathcal M}(z)
  =
  \mathcal J_\Lambda(\rho,U)\,
  \dd^4a\,\dd\rho\,\dd\Omega(U).
\]
Let
\[
  \mathcal P
  =
  \sum_{A=1}^{2}p_{R,A}
  -
  \sum_{B=1}^{2}p_{L,B}.
\]
For plane-wave external slots, write the individual amputated zero-mode
overlaps in the form
\[
  R^{\rm amp}_{Af}(a,\rho,U)
  =
  \rho^{3/2}\ee^{\ii p_{R,A}\cdot a}
  \mathcal R_{Af}(\rho,U;p_{R,A}),
  \qquad
  L^{\rm amp}_{fB}(a,\rho,U)
  =
  \rho^{3/2}\ee^{-\ii p_{L,B}\cdot a}
  \mathcal L_{fB}(\rho,U;p_{L,B}),
\]
where \(\mathcal R\) and \(\mathcal L\) are obtained from the Fourier
transforms of the individual normalized zero-mode wavefunctions, with the
chosen spin-color external covectors already contracted.  Then the
source-differentiated zero-mode coefficient in
\eqref{eq:instanton-amputated-four-fermion-kernel} has the factorized form
\begin{equation}
\begin{aligned}
  \det R^{\rm amp}(a,\rho,U)\det L^{\rm amp}(a,\rho,U)
  &=
  \rho^6\,
  \ee^{\ii\mathcal P\cdot a}\,
  \mathfrak A_2^{\rm lin}(\rho,U;\{p_R,p_L\}),\\
  \mathfrak A_2^{\rm lin}
  &=
  \det\mathcal R(\rho,U;\{p_R\})\,
  \det\mathcal L(\rho,U;\{p_L\}).
\end{aligned}
  \label{eq:instanton-plane-wave-source-determinant}
\end{equation}
A separate bilinear-source specialization is obtained by inserting
\eqref{eq:instanton-amputated-source-matrix} into \(\det B^{\rm amp}\).  In
that case \(F(z)=zK_1(z)\) appears through the density form factor and
\[
  \det B^{\rm amp}
  =
  \rho^6\ee^{\ii\mathcal P\cdot a}
  \left[
    F(\rho|q_{11}|)F(\rho|q_{22}|)
    \mathcal T_{11}\mathcal T_{22}
    -
    F(\rho|q_{12}|)F(\rho|q_{21}|)
    \mathcal T_{12}\mathcal T_{21}
  \right],
  \qquad q_{ff'}=p_{R,f}-p_{L,f'} ,
\]
provided \(B^{\rm amp}\) is an independent bilinear source matrix.  If
\(B^{\rm amp}_{ff'}=u_fv_{f'}\) is only a c-number outer product, the
two-flavor determinant is zero.
Consequently the center integral in infinite volume gives the distributional
factor
\begin{equation}
  \int_{\mathbb R^4}\dd^4a\,
  \ee^{\ii\mathcal P\cdot a}
  =
  (2\pi)^4\delta^{(4)}(\mathcal P),
  \label{eq:instanton-center-momentum-conservation}
\end{equation}
and the source-amputated one-instanton contribution may be written as
\begin{equation}
\begin{aligned}
  \mathcal M^{(4),{\rm amp}}_{\Lambda,Q=1}
  &=
  \ee^{\ii\theta}
  (2\pi)^4\delta^{(4)}(\mathcal P)
  \int
  \exp[-S_\Lambda(A_{\rho,U})]\,
  \mathcal W_{\Lambda,{\rm pt}}^{\rm nz}(\rho,U;\mu) \\
  &\qquad\qquad\qquad
  {}\times
  \rho^6\,
  \mathfrak A_2^{\rm lin}(\rho,U;\{p_R,p_L\})\,
  \mathcal J_\Lambda(\rho,U)\,
  \dd\rho\,\dd\Omega(U).
\end{aligned}
  \label{eq:instanton-plane-wave-four-fermion-amplitude}
\end{equation}
The anti-instanton term has the conjugate chirality structure and the phase
\(\ee^{-\ii\theta}\).

Thus the instanton calculation does not first create a local vertex and then
decorate it.  The finite-regulator amplitude is assembled from four
load-bearing pieces: the center integral, the individual zero-mode
Fourier-overlap tensors, the nonzero-mode determinant, and the size integral.
The local 't Hooft operator is obtained only after the external slots have
been fused into a local bilinear source, the support of the \(\rho\)-integral
lies in the small-momentum regime of the relevant zero-mode form factors, and
the determinant remainder has been replaced by the finite constant of the
chosen scheme.
\end{proposition}

\begin{proof}
In \(\det R^{\rm amp}\), both permutation terms contain the same phase
\(\ee^{\ii(p_{R,1}+p_{R,2})\cdot a}\).  In \(\det L^{\rm amp}\), both
permutation terms contain the same phase
\(\ee^{-\ii(p_{L,1}+p_{L,2})\cdot a}\).  The product therefore carries
\(\ee^{\ii\mathcal P\cdot a}\).  The four individual zero-mode overlaps carry
the conventional total power \(\rho^6\), leaving
\(\det\mathcal R\,\det\mathcal L\) as the nontrivial form-factor and
orientation coefficient.  This proves
\eqref{eq:instanton-plane-wave-source-determinant}.  Translation invariance of
the center coordinate then gives
\eqref{eq:instanton-center-momentum-conservation} in infinite volume, or the
corresponding Kronecker delta on a finite periodic regulator.  Substituting the
factorized coefficient and collective-coordinate density into
\eqref{eq:instanton-amputated-four-fermion-kernel} gives
\eqref{eq:instanton-plane-wave-four-fermion-amplitude}.  The bilinear-source
formula follows from the same phase count applied to
\eqref{eq:instanton-amputated-source-matrix}; the final rank-one warning is
the elementary identity \(\det(u v^T)=0\) for two-component c-number vectors.
\end{proof}

\begin{controlledapproximation}[Hard wave-packet support in the four-slot
instanton kernel]
\label{ca:instanton-hard-wave-packet-source-support}
The plane-wave formula above is a useful coordinate, but a physical hard
source is a wave packet or a source smearing.  In the source-amputated
\(N_f=2\) setup, replace the plane-wave slot covectors by wave packets and
write the reduced zero-mode overlap matrices, after extracting the explicit
\(\rho^{3/2}\) from each slot, as
\begin{equation}
\begin{aligned}
  \mathcal R_{Af}(\rho,U)
  &=
  \int \widehat f^R_A(p)\,
  F_{\rm zm}(\rho |p|/2)\,
  \mathcal T^R_{Af}(\widehat p,U)\,\dd\Pi_R(p),\\
  \mathcal L_{fB}(\rho,U)
  &=
  \int
  F_{\rm zm}(\rho |p|/2)\,
  \mathcal T^L_{fB}(\widehat p,U)\,
  \widehat f^L_B(p)\,\dd\Pi_L(p),
\end{aligned}
  \label{eq:instanton-wave-packet-zero-mode-overlaps}
\end{equation}
where \(\dd\Pi_{R,L}\) denote the chosen finite-regulator momentum measures
and \(\mathcal T^{R,L}\) contain the spin-color tensors and source
normalizations.  Suppose each slot is supported in a hard annulus
\[
  c_-Q\le |p|\le c_+Q,\qquad c_->0,
\]
and suppose the source tensors and wave packets give row-norm bounds
\begin{equation}
  \sum_f|\mathcal R_{Af}(\rho,U)|^2\le
  C^2_{R,A}(1+c_-Q\rho)^{-6},
  \qquad
  \sum_f|\mathcal L_{fB}(\rho,U)|^2\le
  C^2_{L,B}(1+c_-Q\rho)^{-6}.
  \label{eq:instanton-wave-packet-row-bound}
\end{equation}
Such a bound follows, for example, from the large-slot estimate
\(|F_{\rm zm}(t)|\le C_{\rm zm}(1+t)^{-3}\) together with bounded source
norms.  Hadamard's inequality then gives the pointwise four-slot estimate
\begin{equation}
  \left|
  \rho^6\det\mathcal R(\rho,U)\det\mathcal L(\rho,U)
  \right|
  \le
  \rho^6 C_{\rm src}(1+c_-Q\rho)^{-12},
  \qquad
  C_{\rm src}:=\prod_A C_{R,A}\prod_B C_{L,B}.
  \label{eq:instanton-wave-packet-four-slot-bound}
\end{equation}
For the one-loop \(SU(3)\), \(N_f=2\) hard density, this gives the explicit
tail bound
\begin{equation}
\begin{aligned}
  Q^{b_0}
  \int_{R/Q}^{\infty}
  \rho^{b_0+1}
  C_{\rm src}(1+c_-Q\rho)^{-12}\,\dd\rho
  &\le
  Q^{-2}C_{\rm src}c_-^{-12}
  \frac{R^{\,b_0-10}}{10-b_0}  \\
  &=
  3Q^{-2}C_{\rm src}c_-^{-12}R^{-1/3},
\end{aligned}
  \label{eq:instanton-wave-packet-hard-tail-bound}
\end{equation}
where \(b_0=29/3\) and the inequality is written for \(R\) in the region
where \(c_-R\ge1\).  This is a source-level version of the hard-tail estimate
in ControlledApproximation~\ref{ca:hard-size-dominance-vs-convergence}.

The same formula also records two failure modes.  First, the source must
remain rank-two in each chirality after projection to the instanton
zero-mode subspace: if the two right-handed rows of \(\mathcal R\), or the
two left-handed columns of \(\mathcal L\), become proportional, then
\(\det\mathcal R\det\mathcal L=0\) and the selected four-slot channel has no
leading zero-mode coefficient.  Second, a wave packet with support reaching
soft momentum has \(c_-=0\), so
\eqref{eq:instanton-wave-packet-hard-tail-bound} is unavailable.  The
calculation must then use the actual soft source profile, a fused bilinear
source, physical screening, or a Wilsonian long-size remainder.  Hardness is
therefore a property of the source projection and endpoint estimate, not of
the BPST moduli coordinates alone.
\end{controlledapproximation}

\begin{proposition}[Orientation projection in the instanton four-slot kernel]
\label{prop:instanton-color-orientation-haar-projection}
The global \(SU(2)\) orientation in the embedded BPST core enters the
amplitude as a projection of colored zero-mode tensors onto invariant color
tensors.  Normalize Haar measure by \(\int_{SU(2)}\dd U=1\), and let
\(U_i{}^\alpha\) denote the fundamental matrix element carrying a core
two-dimensional index \(\alpha\) to a color index \(i\).  Then
\begin{equation}
  \int_{SU(2)} U_i{}^\alpha\,\dd U=0,\qquad
  \int_{SU(2)}
    U_i{}^\alpha \overline{U_j{}^\beta}\,\dd U
  =
  \frac12\delta_{ij}\delta^{\alpha\beta},
  \label{eq:instanton-orientation-one-slot-projection}
\end{equation}
and, because \(\wedge^2\mathbb C^2\) is the trivial representation of
\(SU(2)\),
\begin{equation}
  \int_{SU(2)}
    U_i{}^\alpha U_j{}^\beta\,\dd U
  =
  \frac12\varepsilon_{ij}\varepsilon^{\alpha\beta}.
  \label{eq:instanton-orientation-two-slot-projection}
\end{equation}
Consequently two differentiated external zero-mode slots with core source
vectors \(v_A,v_B\in\mathbb C^2\) have orientation average
\begin{equation}
  \int_{SU(2)}
    (Uv_A)_i(Uv_B)_j\,\dd U
  =
  \frac12\varepsilon_{ij}\,
  \varepsilon_{\alpha\beta}v_A^\alpha v_B^\beta
  =
  \frac12\varepsilon_{ij}\det(v_A,v_B).
  \label{eq:instanton-orientation-two-slot-determinant}
\end{equation}
For an \(N_f=2\) differentiated four-slot source, the four slots share the same
orientation variable.  One must therefore use the single four-point Haar
projector, not the product of two two-point averages.  When all four selected
slots have been written as fundamental \(SU(2)\) core indices, set
\[
  A_{i_1i_2j_1j_2}
  =
  \varepsilon_{i_1i_2}\varepsilon_{j_1j_2},
  \qquad
  B_{i_1i_2j_1j_2}
  =
  \varepsilon_{i_1j_1}\varepsilon_{i_2j_2},
\]
and use the same symbols with upper core indices.  Then
\begin{equation}
\begin{aligned}
  &\int_{SU(2)}
    U_{i_1}{}^{\alpha_1}U_{i_2}{}^{\alpha_2}
    U_{j_1}{}^{\beta_1}U_{j_2}{}^{\beta_2}\,\dd U       \\
  &\quad =
  \frac13 A_{i_1i_2j_1j_2}A^{\alpha_1\alpha_2\beta_1\beta_2}
  -\frac16 A_{i_1i_2j_1j_2}B^{\alpha_1\alpha_2\beta_1\beta_2} \\
  &\qquad
  -\frac16 B_{i_1i_2j_1j_2}A^{\alpha_1\alpha_2\beta_1\beta_2}
  +\frac13 B_{i_1i_2j_1j_2}B^{\alpha_1\alpha_2\beta_1\beta_2}.
\end{aligned}
  \label{eq:instanton-four-slot-orientation-projector}
\end{equation}
If the left slots are kept as conjugate or dual slots instead, they must first
be converted to the chosen fundamental convention by the \(SU(2)\) pseudoreal
\(\varepsilon\)-tensor, or else the corresponding mixed \(U\overline U\)
Weingarten projector must be used.  Contracting
\eqref{eq:instanton-four-slot-orientation-projector} with core source vectors
\(v_1^R,v_2^R,v_1^L,v_2^L\) gives
\begin{equation}
\begin{aligned}
  &A_{i_1i_2j_1j_2}
  \left[
    \frac13\Delta_{12|34}
    -\frac16\Delta_{13|24}
  \right]
  +
  B_{i_1i_2j_1j_2}
  \left[
    -\frac16\Delta_{12|34}
    +\frac13\Delta_{13|24}
  \right],
\end{aligned}
  \label{eq:instanton-four-slot-source-projector}
\end{equation}
where
\[
  \Delta_{12|34}
  =
  \varepsilon_{\alpha_1\alpha_2}v_1^{R,\alpha_1}v_2^{R,\alpha_2}\,
  \varepsilon_{\beta_1\beta_2}v_1^{L,\beta_1}v_2^{L,\beta_2},
  \qquad
  \Delta_{13|24}
  =
  \varepsilon_{\alpha_1\beta_1}v_1^{R,\alpha_1}v_1^{L,\beta_1}\,
  \varepsilon_{\alpha_2\beta_2}v_2^{R,\alpha_2}v_2^{L,\beta_2}.
\]
Thus the color-orientation collective coordinate is a genuine part of the
amplitude rather than a passive coordinate volume: it enforces the full
invariant four-slot projector.  A product of right and left flavor
determinants appears only after the Berezin source coefficient and the final
color-singlet source contraction impose that additional structure.  For an
\(SU(N_c)\) instanton
embedded through an \(SU(2)\) subgroup, the full orientation-orbit integral
is the corresponding \(SU(N_c)\)-invariant projection with a convention
dependent normalization; the displayed \(SU(2)\) core identity is the local
finite projection used by the zero-mode tensor.
\end{proposition}

\begin{proof}
There is no nonzero invariant vector in the fundamental representation, so
the one-slot average vanishes.  The mixed two-point average is an intertwiner
\(\mathbb C^2\otimes(\mathbb C^2)^*\to
\mathbb C^2\otimes(\mathbb C^2)^*\), hence Schur orthogonality makes it a
multiple of \(\delta_{ij}\delta^{\alpha\beta}\).  Contracting \(i=j\) and
\(\alpha=\beta\) gives \(2\), so the multiple is \(1/2\).

Similarly, the average of \(U_i{}^\alpha U_j{}^\beta\) is an invariant tensor
in \(\wedge^2\mathbb C^2\otimes\wedge^2(\mathbb C^2)^*\), and must be
\(c\,\varepsilon_{ij}\varepsilon^{\alpha\beta}\).  Contracting with
\(\varepsilon^{ij}\varepsilon_{\alpha\beta}\) gives
\[
  4c
  =
  \int
  \varepsilon^{ij}\varepsilon_{\alpha\beta}
  U_i{}^\alpha U_j{}^\beta\,\dd U
  =
  2\int\det U\,\dd U
  =
  2,
\]
so \(c=1/2\).  Multiplying the averaged matrix elements by the two source
vectors gives
\eqref{eq:instanton-orientation-two-slot-determinant}.

For four slots, the invariant subspace of
\((\mathbb C^2)^{\otimes4}\) is two-dimensional.  Use the two spanning
invariant tensors \(A\) and \(B\) above.  Their Gram matrix is
\[
  \begin{pmatrix}
  \langle A,A\rangle & \langle A,B\rangle\\
  \langle B,A\rangle & \langle B,B\rangle
  \end{pmatrix}
  =
  \begin{pmatrix}4&2\\2&4\end{pmatrix},
  \qquad
  G^{-1}
  =
  \begin{pmatrix}\frac13&-\frac16\\[2pt]-\frac16&\frac13\end{pmatrix}.
\]
The Haar average is the orthogonal projection onto this invariant subspace,
with one copy of \(G^{-1}\) connecting the color invariant basis to the core
invariant basis.  This gives
\eqref{eq:instanton-four-slot-orientation-projector}.  Contracting that
projector with the four core source vectors gives
\eqref{eq:instanton-four-slot-source-projector}.  The calculation uses one
shared Haar variable throughout; replacing it by a product of two two-slot
averages would project two independent orientations and gives the wrong
four-slot normalization.
\end{proof}

The qualification in
Proposition~\ref{prop:instanton-external-leg-amputation-kernel} is essential.
Colored quark legs in confining QCD do not define ordinary physical
asymptotic states.  The displayed object is an auxiliary source-amputated, or
partonic hard-kernel, coordinate useful for matching the finite-regulator
Green function to the local effective operator.  Gauge-invariant amplitudes
must contract the same instanton kernel with color-singlet external
operators, hadronic matrix elements, or a separately specified infrared
deformation.  This is the amplitude-level content of the original
one-instanton computation of 't Hooft
\cite{tHooft1976Pseudoparticle}: collective coordinates, nonzero-mode
fluctuation determinants, fermion zero-mode saturation, and external-leg
normalization are one calculation.

\begin{controlledapproximation}[Color-singlet matching of the instanton
kernel]
\label{ca:instanton-color-singlet-hadronic-matching}
Stay in the finite-regulator \(Q=1\) setup of
Proposition~\ref{prop:instanton-external-leg-amputation-kernel}, but replace
the auxiliary colored external quark representatives by renormalized
gauge-invariant source families.  Local examples are chiral color-singlet
bilinears \(\bar q_{R f}\Gamma q_{L f'}\); nonlocal examples must include the
Wilson line which makes the two quark fields a color singlet.  In a selected
\(N_f=2\) chirality-violating channel, suppose that the zero-mode projection
of the chosen source family has the short-distance form
\begin{equation}
  B^{\mathcal J}_{AB}(z)
  =
  \sum_{f,f'=1}^{2}
  \Phi^{R}_{A f}(z;\mu)\,
  B^{\rm amp}_{ff'}(z)\,
  \Phi^{L}_{f' B}(z;\mu)
  +
  B^{\rm rem}_{AB}(z),
  \label{eq:instanton-color-singlet-source-projection}
\end{equation}
where \(A,B\) label two independent color-singlet source coordinates,
\(\Phi^R,\Phi^L\) are the source-to-zero-mode overlap matrices in the chosen
renormalization scheme, and \(B^{\rm rem}\) collects higher-dimension source
components, soft channels not represented by the hard zero-mode kernel, and
terms outside the selected chirality structure.  The corresponding
gauge-invariant source correlator receives the one-instanton contribution
\begin{equation}
  G^{(4)}_{\mathcal J,\Lambda,Q=1}
  =
  \int_{\mathcal U_{Q=1}}
  \exp[-S_\Lambda(A_z)+\ii\theta]\,
  \mathcal W_{\Lambda,{\rm pt}}^{\rm nz}(z;\mu)\,
  \det B^{\mathcal J}(z)\,
  \dd\nu_{\Lambda,\mathcal M}(z),
  \label{eq:instanton-color-singlet-source-correlator}
\end{equation}
up to the same remainder specification.  If the overlap matrices are square
and vary only on scales which are soft compared with the hard instanton
support, then pointwise in \(z\)
\begin{equation}
  \det B^{\mathcal J}(z)
  =
  \det\Phi^R(z;\mu)\,
  \det B^{\rm amp}(z)\,
  \det\Phi^L(z;\mu)
  +
  \hbox{terms containing }B^{\rm rem}.
  \label{eq:instanton-color-singlet-determinant-projection}
\end{equation}
Thus the zero-mode form factors appropriate to the selected source family, the
orientation tensor, the nonzero-mode determinant, and the size integral remain
the short-distance instanton kernel; the color-singlet source overlaps are an
additional matching layer, not a replacement for the fluctuation calculation.

For stable hadron external states, the physical amplitude is obtained only
after applying the usual color-singlet spectral or LSZ extraction to the
gauge-invariant correlator.  For example, if interpolating sources
\(\mathcal I_H,\mathcal I_{H'}\) have nonzero isolated overlaps
\(Z_H,Z_{H'}\) with stable finite-volume states in the sense of
Proposition~\ref{prop:qcd-stable-current-three-point-extraction}, then a
termwise semiclassical extraction requires the same pole-residue operation:
\begin{equation}
  \mathcal A^{(Q=1)}_{H'\leftarrow H}
  =
  \bigl(\overline Z_{H'}Z_H\bigr)^{-1}
  \operatorname*{Res}_{H',H}
  G_{\mathcal I_{H'}^\dagger\,\mathcal X\,\mathcal I_H,\Lambda,Q=1},
  \label{eq:instanton-hadronic-residue-extraction}
\end{equation}
where \(\mathcal X\) is the chosen gauge-invariant current or source product
and the residue notation means the finite-volume large-time projection, or
the corresponding infinite-volume pole extraction when it exists.  The BPST
calculation does not determine \(Z_H,Z_{H'}\), the hadronic matrix element of
the induced 't Hooft operator, or the long-distance part of
\(\Phi^{R,L}\).  Those are nonperturbative QCD data.  In a hard OPE or
factorization regime, the small-\(\rho\) instanton calculation can supply a
Wilson coefficient multiplying a gauge-invariant operator such as
\(\det[\bar q_R q_L]_\mu\); outside such a regime
\eqref{eq:instanton-color-singlet-source-correlator} is a source correlator,
not a prediction for a hadronic amplitude by itself.
\end{controlledapproximation}

The finite-regulator bookkeeping behind
\eqref{eq:instanton-color-singlet-source-projection} is entirely
finite-dimensional.
At finite regulator the fermion fields have finitely many odd coefficients,
and every regulated source has a finite expansion in those coefficients.
Restricting the source to the instanton zero-mode subspace produces a
finite-dimensional polynomial in the zero-mode variables.  The hypothesis
\eqref{eq:instanton-color-singlet-source-projection} says that, in the
chosen hard channel, the part of this polynomial which is bilinear in the
right and left zero-mode variables factors through the same amputated
zero-mode kernel \(B^{\rm amp}\), with source-dependent overlap maps on the
two sides.  Substituting this zero-mode source matrix into the Berezin
coefficient in
Proposition~\ref{prop:qcd-thooft-determinant-zero-mode-amplitude} gives
\(\det B^{\mathcal J}(z)\), and multiplying by the determinant density and
collective-coordinate measure gives
\eqref{eq:instanton-color-singlet-source-correlator}.

The determinant relation
\eqref{eq:instanton-color-singlet-determinant-projection} is the finite
identity
\[
  \det(\Phi^R B^{\rm amp}\Phi^L)
  =
  \det\Phi^R\,\det B^{\rm amp}\,\det\Phi^L,
\]
applied pointwise in the collective coordinate.  If the overlaps vary with
\(z\), the same identity still holds pointwise but the overlap factors must
remain inside the collective-coordinate integral.  Finally, because
\(G_{\mathcal I_{H'}^\dagger\,\mathcal X\,\mathcal I_H,\Lambda,Q=1}\) is a
correlator of gauge-invariant sources, the hadronic extraction is the
standard spectral-residue operation for color-singlet correlators.  Applying
that operation termwise in a controlled semiclassical expansion gives
\eqref{eq:instanton-hadronic-residue-extraction}; it never produces an
ordinary colored-quark \(S\)-matrix element.

\begin{controlledapproximation}[Euclidean instanton kernel to physical
amplitude]
\label{ca:instanton-euclidean-to-physical-amplitude-budget}
Let \(G^{E,Q=1}_{\mathcal J,\Lambda}\) be the finite-regulator
gauge-invariant Euclidean source correlator obtained from the one-instanton
chart after zero modes, nonzero-mode determinants, source projections, and
collective coordinates have all been paired.  Let
\[
  \mathcal A_{\rm inst}^{\rm lead}
  =
  \operatorname*{Proj}_{\rm phys}
  \operatorname*{Cont}_{\mathcal S}
  G^{E,Q=1}_{\mathcal J,{\rm lead}}
\]
denote the result of first replacing the finite-regulator density by the
declared leading instanton kernel, then analytically continuing on a specified
sheet \(\mathcal S\), and finally applying the chosen physical projection:
a stable-particle pole residue, an inclusive spectral weight, or a hard OPE
coefficient.  The corresponding physical one-instanton contribution
\(\mathcal A_{\rm phys}^{Q=1}\) differs from this coordinate by
\begin{equation}
  \mathcal A_{\rm phys}^{Q=1}
  -
  \mathcal A_{\rm inst}^{\rm lead}
  =
  R_{\rm reg}
  +
  R_{\rm cont}
  +
  R_{\rm spec}
  +
  R_{\rm IR}
  +
  R_{\rm cut}
  +
  R_{\rm match}
  +
  R_{\rho}.
  \label{eq:instanton-euclidean-to-physical-residual-budget}
\end{equation}
Here \(R_{\rm reg}\) records finite-volume, cutoff, and order-of-limits errors
between the regulated Euclidean sector and its continuum source correlator;
\(R_{\rm cont}\) records analytic-continuation and sheet choices, including
thresholds and singularities crossed by the chosen continuation;
\(R_{\rm spec}\) records the pole, spectral, or OPE projection error;
\(R_{\rm IR}\) records confinement, massless radiation, inclusive-sum, or
external Wilson-line data not contained in a colored partonic kernel;
\(R_{\rm cut}\) records unitarity-cut and conjugate-sector effects, including
instanton--anti-instanton sectors needed before an \(S\)-matrix statement can
obey an optical theorem; \(R_{\rm match}\) records the matching from the
chosen source family to the local operator or hadronic observable; and
\(R_{\rho}\) is the remaining instanton-size endpoint and determinant-tail
contribution after the selected hard, screened, or Wilsonian window has been
declared.

If on the selected kinematic window each residual satisfies
\[
  |R_\kappa|\le \varepsilon_\kappa,
  \qquad
  \kappa\in
  \{{\rm reg},{\rm cont},{\rm spec},{\rm IR},{\rm cut},{\rm match},\rho\},
\]
then
\begin{equation}
  \left|
  \mathcal A_{\rm phys}^{Q=1}
  -
  \mathcal A_{\rm inst}^{\rm lead}
  \right|
  \le
  \sum_\kappa\varepsilon_\kappa .
  \label{eq:instanton-physical-amplitude-residual-bound}
\end{equation}
The point is not to add another approximation.  It is to identify where the
physics lives after the moduli-space and determinant calculation has been
done.  In a weakly coupled Higgsed theory, a thermal screened regime, or a
hard Euclidean OPE window, the physical projection and endpoint controls can
sometimes be declared explicitly.  In confining zero-temperature QCD, a
colored four-slot kernel is only a short-distance source coordinate until the
gauge-invariant spectral or OPE projection and the long-size remainder have
been supplied.
\end{controlledapproximation}

\begin{controlledapproximation}[Instanton cut pairing for inclusive rates]
\label{ca:instanton-unitarity-cut-pairing}
Stay in the source channel of
ControlledApproximation~\ref{ca:instanton-euclidean-to-physical-amplitude-budget}.
Before making a rate statement, separate the amplitude into the sectors which
can interfere on the chosen cut:
\begin{equation}
  \mathcal A_{\rm phys}[\mathcal J]
  =
  A_0[\mathcal J]
  +
  \ee^{-S_I+\ii\theta}a_I[\mathcal J]
  +
  \ee^{-S_I-\ii\theta}a_{\bar I}[\mathcal J]
  +
  \cdots .
  \label{eq:instanton-amplitude-sector-expansion}
\end{equation}
For a finite-regulator inclusive cut pairing \(\langle\,,\,\rangle_{\rm cut}\),
the first terms in the corresponding rate are therefore
\begin{equation}
\begin{aligned}
  \Gamma[\mathcal J]
  &=
  \langle A_0,A_0\rangle_{\rm cut}
  +
  2\operatorname{Re}\!
  \left[
    \ee^{-S_I+\ii\theta}
    \langle A_0,a_I\rangle_{\rm cut}
  \right] \\
  &\quad
  +
  2\operatorname{Re}\!
  \left[
    \ee^{-S_I-\ii\theta}
    \langle A_0,a_{\bar I}\rangle_{\rm cut}
  \right]
  +
  \ee^{-2S_I}
  \langle a_I,a_I\rangle_{\rm cut}
  +
  \cdots .
\end{aligned}
  \label{eq:instanton-inclusive-rate-sector-expansion}
\end{equation}
Thus an order-\(\ee^{-S_I}\) interference term is present unless the selected
source and final-state projection kill it.  In the chirality-violating
't Hooft channel this can happen because the perturbative \(Q=0\) sector has no
zero modes to supply the selected axial-charge violation, but this is a
selection statement supplied by the source projection and final-state cut.
The hypothesis needed below is
\begin{equation}
  \langle B,a_I\rangle_{\rm cut}
  =
  \langle B,a_{\bar I}\rangle_{\rm cut}
  =
  0
  \quad
  \text{for every lower-action sector }B\text{ retained by the projection}.
  \label{eq:instanton-cut-selection-hypothesis}
\end{equation}

After the determinant, zero-mode, orientation, and source-matching factors
have been assembled, write a finite one-instanton amplitude coordinate as
\begin{equation}
  A_I[\mathcal J]
  =
  \ee^{\ii\theta}
  \int_{\mathcal U_I}
  K_I(z;\mathcal J)\,\dd\nu_I(z),
  \label{eq:instanton-one-amplitude-kernel-coordinate}
\end{equation}
where \(z\) denotes the collective coordinates and finite fluctuation/source
data retained in the chosen regulator.  The conjugate channel has an
anti-instanton density
\[
  A_{\bar I}[\mathcal J^\dagger]
  =
  \ee^{-\ii\theta}
  \int_{\mathcal U_{\bar I}}
  K_{\bar I}(\bar z;\mathcal J^\dagger)\,\dd\nu_{\bar I}(\bar z).
\]
The factorized diagonal instanton contribution to the selected inclusive
spectral weight or cut rate has the paired form
\begin{equation}
  \Gamma^{(0),{\rm diag}}_{I\bar I}[\mathcal J]
  =
  \int_{\mathcal U_I\times\mathcal U_{\bar I}}
  K_I(z;\mathcal J)\,
  K_{\bar I}(\bar z;\mathcal J^\dagger)\,
  \mathcal C_{\rm cut}^{(0)}(z,\bar z)\,
  \dd\nu_I(z)\,\dd\nu_{\bar I}(\bar z).
  \label{eq:instanton-cut-paired-rate-leading}
\end{equation}
Here \(\mathcal C_{\rm cut}^{(0)}\) contains the chosen leading final-state
sum, on-shell or spectral projector, relative \(I\bar I\) separation and
orientation weight, and leading relative fluctuation determinant.  At the
factorized diagonal level the topological phases cancel in the cut channel,
\[
  \ee^{\ii\theta}\ee^{-\ii\theta}=1,
  \qquad
  \ee^{\ii\theta}\ee^{\ii\theta}=\ee^{2\ii\theta}.
\]
Thus a same-charge pair belongs to a charge-two amplitude sector rather than
to the amplitude-times-conjugate contribution to a neutral inclusive rate.  The
displayed kernel is the diagonal \(I\bar I\) contribution to the selected
inclusive rate.  It is the leading nonperturbative contribution to the full
selected rate, with semiclassical exponential \(e^{-2S_I}\) and no explicit
\(\theta\) phase, only when
\eqref{eq:instanton-cut-selection-hypothesis} has removed the lower-order
interference terms in \eqref{eq:instanton-inclusive-rate-sector-expansion}.
Without that hypothesis, the order-\(\ee^{-S_I}\) interference is part of the
physical rate and can carry explicit \(\theta\)-dependence.

The full cut kernel can be organized as
\[
  \mathcal C_{\rm cut}
  =
  \mathcal C_{\rm cut}^{(0)}(1+\delta_{\rm cut}),
  \qquad
  |\delta_{\rm cut}(z,\bar z)|\le \varepsilon_{\rm cut},
\]
where \(\delta_{\rm cut}\) absorbs the difference between the declared leading
cut package and the actual final-state sum, relative \(I\bar I\) interaction,
valley or quasi-zero-mode treatment, and nonzero-mode determinant ratio.  If
\[
  M^{(0)}_{I\bar I}
  =
  \int
  |K_I(z;\mathcal J)|\,|K_{\bar I}(\bar z;\mathcal J^\dagger)|\,
  |\mathcal C_{\rm cut}^{(0)}(z,\bar z)|\,
  \dd\nu_I(z)\,\dd\nu_{\bar I}(\bar z)
  <\infty ,
\]
then the exact cut contribution satisfies
\begin{equation}
  \left|
  \Gamma_{I\bar I}[\mathcal J]
  -
  \Gamma^{(0),{\rm diag}}_{I\bar I}[\mathcal J]
  \right|
  \le
  \varepsilon_{\rm cut} M^{(0)}_{I\bar I}.
  \label{eq:instanton-cut-pairing-error-bound}
\end{equation}
This is the concrete finite-regulator content of the \(R_{\rm cut}\) entry in
\eqref{eq:instanton-euclidean-to-physical-residual-budget}.  The local
't Hooft vertex supplies the chirality-violating amplitude kernel.  Inclusive
cross sections, decay rates, and spectral weights require the conjugate
anti-instanton density, final-state cut, and relative \(I\bar I\)
fluctuation determinants and interactions in addition to that kernel.
\end{controlledapproximation}

\begin{controlledapproximation}[One-instanton sector isolation bound]
\label{ca:one-instanton-sector-isolation-bound}
The previous cut-pairing discussion separates amplitude sectors for rates.
For a selected one-instanton amplitude there is a prior, more elementary
question: has the \(Q=1\) source coordinate actually been isolated from the
neighboring sectors in the observable under study?  Let \(\Pi_{\mathcal O}\)
denote the finite source, charge, chirality, or spectral projection used to
extract the coefficient.  On the same regulator and kinematic window write
the projected amplitude-sector expansion as
\begin{equation}
  \Pi_{\mathcal O}\mathcal A
  =
  \Pi_{\mathcal O}A_0
  +
  \Pi_{\mathcal O}A_I
  +
  \Pi_{\mathcal O}A_{\bar I}
  +
  \Pi_{\mathcal O}A_{2I}
  +
  \Pi_{\mathcal O}A_{I\bar I}^{\rm amp}
  +
  R_{\ge2}^{\rm other}.
  \label{eq:instanton-one-sector-isolation-expansion}
\end{equation}
Here \(A_{I\bar I}^{\rm amp}\) is an amplitude-sector contribution before
forming an inclusive cut rate; it is not the diagonal rate term in
\eqref{eq:instanton-cut-paired-rate-leading}.  Let \(J_I\) be the retained
leading \(Q=1\) coefficient, obtained after the determinant, zero-mode,
orientation, source, and size-window choices have all been made, and let
\(M_I\) be the absolute mass of that retained \(Q=1\) window.  Suppose
\begin{equation}
  \left|\Pi_{\mathcal O}A_I-J_I\right|
  \le
  \epsilon_I M_I,
  \qquad
  \left|\Pi_{\mathcal O}A_\alpha\right|
  \le
  \lambda_\alpha M_\alpha
  \quad
  (\alpha=0,\bar I,2I,I\bar I),
  \qquad
  |R_{\ge2}^{\rm other}|\le E_{\ge2}.
  \label{eq:instanton-one-sector-isolation-hypotheses}
\end{equation}
Then the selected observable differs from the retained one-instanton
coefficient by
\begin{equation}
  \left|\Pi_{\mathcal O}\mathcal A-J_I\right|
  \le
  \epsilon_I M_I
  +
  \lambda_0M_0
  +
  \lambda_{\bar I}M_{\bar I}
  +
  \lambda_{2I}M_{2I}
  +
  \lambda_{I\bar I}M_{I\bar I}
  +
  E_{\ge2}.
  \label{eq:instanton-one-sector-isolation-bound}
\end{equation}
If in addition the retained one-instanton coefficient has a noncancellation
margin
\[
  |J_I|\ge\kappa_I M_I,\qquad \kappa_I>0,
\]
then \eqref{eq:instanton-one-sector-isolation-bound} gives a relative
isolation bound after division by \(|J_I|\).  The projection factors
\(\lambda_\alpha\) are part of the physics, not bookkeeping decoration.  In a
chirality-violating source the perturbative leakage \(\lambda_0M_0\) can
vanish by an exact zero-mode selection rule.  In an ordinary gauge-invariant
spectral channel it usually does not, and a small instanton saddle can be
buried under a lower-action sector.  Likewise, signed cancellation between
\(\bar I\), \(2I\), and \(I\bar I\) contributions is not an isolation
bound; the estimate uses absolute sector majorants in the same source
normalization.  A quoted ``one-instanton amplitude'' is therefore a leading
sector coordinate until the sector projection and leakage bounds above have
also been supplied.
\end{controlledapproximation}

\begin{proposition}[One-instanton amplitude assembly and error budget]
\label{prop:instanton-amplitude-assembly-error-budget}
Fix a finite-regulator \(Q=1\) chart and a gauge-invariant source channel.
Let
\begin{equation}
  \mathcal A_{\Lambda,Q=1}^{\mathcal J}
  =
  \int_{\mathcal U_{Q=1}}
  \mathcal K_{\rm ex}(z)\,\dd z
  \label{eq:instanton-exact-source-amplitude-density}
\end{equation}
denote the exact one-instanton contribution to the selected source
correlator after gauge fixing but before replacing any finite factor by its
semiclassical leading term.  Choose a controlled subdomain
\(\mathcal U_{\rm hard}\subset\mathcal U_{Q=1}\) and a leading density
\begin{equation}
\begin{aligned}
  \mathcal K_{\rm lead}(z)
  &=
  \ee^{\ii\theta}
  \mathcal C_{\rm pg}^{\mathcal S}(z)\,
  \mathcal R_{\rm f}^{\mathcal S}(z;\mu)
  \left(\frac{8\pi^2}{g_{\rm ht}^2(\mu)}\right)^{2N_c}
  \exp\!\left[-\frac{8\pi^2}{g_{\rm ht}^2(\mu)}\right]
  (\mu\rho)^{b_0}\rho^{-5}\,
  \mathcal Z_{\rm lead}^{0}(z).
\end{aligned}
  \label{eq:instanton-leading-amplitude-density}
\end{equation}
Here \(\mathcal C_{\rm pg}^{\mathcal S}(z)\) is the pure-gauge
collective-coordinate and determinant density in the chosen convention.  In
the Pauli-Villars convention after the embedded orientation orbit has been
integrated it contains \(C_{N_c}^{\rm PV}\); in an unintegrated orientation
chart it contains the corresponding orientation density instead of an
already-integrated orbit volume.  The factor
\(\mathcal R_{\rm f}^{\mathcal S}\) is the light-fermion nonzero-mode
determinant factor in the same source scheme, and
\(\mathcal Z_{\rm lead}^{0}\) is the leading zero-mode source coefficient.
Depending on the source type,
\(\mathcal Z_{\rm lead}^{0}\) is one of
\(\det(M^0+J^0)\), \(\det R^{\rm amp}\det L^{\rm amp}\), or the
color-singlet determinant \(\det B^{\mathcal J}\) after the source matching
of \eqref{eq:instanton-color-singlet-source-projection}.

Suppose that on \(\mathcal U_{\rm hard}\) the exact density decomposes as
\begin{equation}
  \mathcal K_{\rm ex}
  =
  \mathcal K_{\rm lead}
  \bigl(1+\varepsilon_{\rm det}
          +\varepsilon_{\rm zm}
          +\varepsilon_{\rm src}\bigr)
  +\Delta_{\rm Schur},
  \label{eq:instanton-amplitude-error-density}
\end{equation}
where \(\varepsilon_{\rm det}\) records the higher-loop and finite determinant
remainder after the chosen determinant convention, \(\varepsilon_{\rm zm}\)
records the zero-mode projection or local-source truncation error,
\(\varepsilon_{\rm src}\) records the color-singlet source-matching or
hadronic/OPE matching remainder, and \(\Delta_{\rm Schur}\) records the
finite nonzero-mode Schur correction from
\eqref{eq:instanton-fermion-schur-factorization}.  Then
\begin{equation}
  \mathcal A_{\Lambda,Q=1}^{\mathcal J}
  =
  \mathcal A_{\rm lead}^{\mathcal J}
  +R_{\rm det}+R_{\rm zm}+R_{\rm src}+R_{\rm Schur}+R_{\rm end},
  \label{eq:instanton-amplitude-error-decomposition}
\end{equation}
with
\[
  \mathcal A_{\rm lead}^{\mathcal J}
  =
  \int_{\mathcal U_{\rm hard}}\mathcal K_{\rm lead}(z)\,\dd z,
  \qquad
  R_{\rm end}
  =
  \int_{\mathcal U_{Q=1}\setminus\mathcal U_{\rm hard}}
  \mathcal K_{\rm ex}(z)\,\dd z .
\]
Moreover, whenever the right-hand side is absolutely integrable,
\begin{equation}
\begin{aligned}
  \bigl|
  \mathcal A_{\Lambda,Q=1}^{\mathcal J}
  -\mathcal A_{\rm lead}^{\mathcal J}
  \bigr|
  &\le
  \int_{\mathcal U_{\rm hard}}
  |\mathcal K_{\rm lead}|
  \bigl(
    |\varepsilon_{\rm det}|
    +|\varepsilon_{\rm zm}|
    +|\varepsilon_{\rm src}|
  \bigr)\,\dd z  \\
  &\quad
  +\int_{\mathcal U_{\rm hard}}|\Delta_{\rm Schur}|\,\dd z
  +
  \int_{\mathcal U_{Q=1}\setminus\mathcal U_{\rm hard}}
  |\mathcal K_{\rm ex}|\,\dd z .
\end{aligned}
  \label{eq:instanton-amplitude-error-bound}
\end{equation}
Thus replacing the finite-regulator instanton contribution by the local
't Hooft operator, or by a hard source-amputated kernel, is justified only
after the determinant remainder, zero-mode/source projection, matching
remainder, Schur correction, and endpoint integral have all been controlled
in the same scheme.
\end{proposition}

\begin{proof}
Split the exact integral into the chosen hard/source domain and its complement:
\[
  \int_{\mathcal U_{Q=1}}\mathcal K_{\rm ex}
  =
  \int_{\mathcal U_{\rm hard}}\mathcal K_{\rm ex}
  +
  \int_{\mathcal U_{Q=1}\setminus\mathcal U_{\rm hard}}\mathcal K_{\rm ex}.
\]
Substituting
\eqref{eq:instanton-amplitude-error-density} on
\(\mathcal U_{\rm hard}\) gives the five remainders in
\eqref{eq:instanton-amplitude-error-decomposition}.  The bound
\eqref{eq:instanton-amplitude-error-bound} is the triangle inequality.  The
content is the identification of the separate physical origins of the five
terms: collective-coordinate geometry supplies only \(\mathcal J_{\rm coll}\)
or, equivalently, the collective part of
\(\mathcal C_{\rm pg}^{\mathcal S}\) and the integration domain; the
fluctuation determinant, light-fermion nonzero-mode factor, zero-mode source
coefficient, source matching, and endpoint behavior are independent
load-bearing inputs to the amplitude.
\end{proof}

Consequently an instanton moduli-space measure by itself is never the
computed QFT amplitude.  It becomes a physical contribution only after it is
multiplied by the nonzero-mode fluctuation determinant, the correctly typed
zero-mode source coefficient, the source or hadron matching data, and an
endpoint estimate for the selected kinematic regime.  The original
one-instanton calculation is deep precisely because it performs this whole
assembly, rather than because it merely parametrizes the BPST collective
coordinates.

\begin{controlledapproximation}[Instanton observable assembly ladder]
\label{ca:instanton-observable-assembly-ladder}
The preceding bounds address different layers of the same physical
calculation.  To prevent the reader from mistaking them for independent
annexes, fix a selected observable channel \(\mathcal O\), a source family
\(\mathfrak h\), and a retained one-instanton window.  Let
\[
  J_I[\mathfrak h]
\]
be the leading finite-window coefficient after the collective-coordinate
density, pure-gauge determinant convention, light-fermion nonzero-mode
factor, source zero-mode determinant, orientation projection, and hard-size
window have all been included.  Let \(M_I[\mathfrak h]\) be its absolute
window mass.  The physical observable map is not the identity map on this
Euclidean source coefficient.  It contains analytic continuation, LSZ or OPE
projection, infrared or hadronic matching, possible inclusive cuts, and the
sector projection used to isolate the \(Q=1\) amplitude coordinate.  Write its
leading linear part as \(\mathfrak L_{\mathcal O}\).

The finite-regulator statement has the form
\begin{equation}
\begin{aligned}
  \mathcal O_{\rm phys}^{(Q=1)}
  &=
  \mathfrak L_{\mathcal O}\!\left[J_I[\mathfrak h]\right]
  +R_{\rm spec}+R_{\rm fluc}+R_{\rm rank}+R_{\rm Schur}
  +R_{\rm end}                                      \\
  &\qquad{}
  +R_{\rm sector}+R_{\rm cont}+R_{\rm proj}
  +R_{\rm IR}+R_{\rm cut}+R_{\rm scheme}.
\end{aligned}
  \label{eq:instanton-observable-assembly-ladder}
\end{equation}
The first line is internal to the Euclidean instanton source integral.  It
records determinant and counterterm remainders, source fluctuation cumulants,
zero-mode rank or source-overlap loss, finite Schur complement corrections,
and the omitted size or boundary tail.  The second line promotes that source
coordinate to the declared physical observable.  It records leakage from
neighboring amplitude sectors, continuation residuals, LSZ/OPE or hadronic
projection errors, infrared and volume residuals, cut errors when a rate is
formed, and finite scheme transport errors.

If these residuals are bounded in the same source normalization by
\[
  |R_\alpha|\le\delta_\alpha M_I[\mathfrak h],
  \qquad
  \alpha\in
  \{\mathrm{spec},\mathrm{fluc},\mathrm{rank},\mathrm{Schur},
    \mathrm{end},\mathrm{sector},\mathrm{cont},\mathrm{proj},
    \mathrm{IR},\mathrm{cut},\mathrm{scheme}\},
\]
then
\begin{equation}
  \left|
  \mathcal O_{\rm phys}^{(Q=1)}
  -
  \mathfrak L_{\mathcal O}[J_I]
  \right|
  \le
  M_I
  \sum_\alpha\delta_\alpha .
  \label{eq:instanton-observable-assembly-bound}
\end{equation}
A relative physical statement further requires a margin
\[
  |\mathfrak L_{\mathcal O}[J_I]|\ge\kappa_{\mathcal O} M_I,
  \qquad \kappa_{\mathcal O}>0.
\]
Only after this margin is present may one divide
\eqref{eq:instanton-observable-assembly-bound} by the quoted leading
observable.  Signed cancellation among the residuals, or between neighboring
amplitude sectors, does not replace the absolute budget.

This ladder is the structural distinction between an instanton-moduli
calculation and an instanton contribution to a QFT observable.  The moduli
space supplies the variables and zero-mode Jacobian inside \(J_I\).  The
fluctuation determinant, source stability, boundary estimates, sector
projection, and physical observable map decide whether \(J_I\) is a
controlled contribution to a scattering amplitude, an OPE coefficient, a
susceptibility, or an inclusive rate.  The operations are deliberately ordered:
one first assembles and bounds the finite source coefficient, then applies
the physical projection and its own residual budget.  Reversing this order
usually replaces a controlled amplitude calculation by a formal local vertex.
\end{controlledapproximation}

\begin{controlledapproximation}[Source-dependent fluctuation bound]
\label{ca:instanton-source-fluctuation-cumulant-bound}
The nonzero-mode determinant in
\eqref{eq:instanton-leading-amplitude-density} is only the Gaussian
normalization of the fluctuation integral.  For an actual source or operator
insertion one must also control how the same nonzero modes enter the insertion.
After zero modes and residual gauge modes have been projected out, let
\(\eta\in\mathcal N_z\) be the finite-regulator normal fluctuation coordinate
at the instanton \(A_z\), let \(d\gamma_z(\eta)\) be the normalized Gaussian
measure defined by the nonzero-mode Hessian, let \(V_z(\eta;g)\) collect cubic
and higher action vertices together with finite counterterm remainders, and let
\(\mathcal Z_z(\eta;\mathcal J)\) be the source factor after the explicit
fermion zero-mode coefficient has been separated.  On a chart where
\(\mathcal Z_z(0;\mathcal J)\ne0\), the source-dependent fluctuation ratio is
\[
  \mathcal R_{\rm fluc}(z;\mathcal J)
  =
  \frac{
    \left\langle
      \mathcal Z_z(\eta;\mathcal J)
      \exp[-V_z(\eta;g)]
    \right\rangle_{\gamma_z}
  }{
    \mathcal Z_z(0;\mathcal J)
  } .
\]
Thus the determinant constant and the source fluctuation correction are
different coordinates: the first normalizes \(d\gamma_z\), while the second
tests the selected source under the same fluctuation measure.

Suppose that on the retained hard/source window
\[
  \mathcal R_{\rm fluc}(z;\mathcal J)
  =
  1+C_1(z;\mathcal J)+C_2(z;\mathcal J)+R_{\ge3}(z;\mathcal J)
\]
and that
\[
  |C_1|\le\epsilon_1,\qquad
  |C_2|\le\epsilon_2,\qquad
  |R_{\ge3}|\le\epsilon_{\ge3}.
\]
Then the fluctuation part of the determinant residual in
\eqref{eq:instanton-amplitude-error-density} is bounded by
\[
  \epsilon_{\rm fluc}
  =
  \epsilon_1+\epsilon_2+\epsilon_{\ge3}.
\]
For the hard finite-window bound this refines
\eqref{eq:hard-instanton-window-absolute-bound} to
\[
  \left|
  C_{\rm inst}^{(4),{\rm ex}}(Q)-\mathcal N(Q)J_R
  \right|
  \le
  |\mathcal N(Q)|M_R
  \bigl(
    \epsilon_{\rm spec}
    +\epsilon_{\rm fluc}
    +\epsilon_{\rm zm}
    +\epsilon_{\rm src}
    +\epsilon_{\rm Schur}
    +\tau_R
  \bigr),
\]
where \(\epsilon_{\rm spec}\) denotes the source-independent spectral
determinant and counterterm remainder.  If
\(\mathcal Z_z(0;\mathcal J)\) loses its rank or noncancellation margin, this
relative form is invalid; the fluctuation contribution must then be kept as an
absolute source residual of the kind represented by \(\Delta_{\rm Schur}\) or
\(R_{\rm src}\).

This is the semiclassical point hidden by many shorthand presentations of the
't Hooft vertex.  A vacuum determinant calibration can fix the Gaussian
normalization and still say too little about a four-fermion amplitude, because
the latter asks whether the source insertion is stable under fluctuations
around the instanton.  The controlled weak-coupling claim is therefore not
``BPST moduli space plus a determinant constant.''  It is a uniform statement,
on the chosen source and size window, about the determinant, zero-mode
overlaps, fluctuation cumulants, endpoint tails, and physical projection
residuals together.
\end{controlledapproximation}

\begin{controlledapproximation}[Hard-momentum size window]
\label{ca:hard-momentum-instanton-size-window}
Consider the \(N_f=2\) source-amputated four-fermion kernel
\eqref{eq:instanton-amputated-four-fermion-kernel} in a kinematic region
where the hard arguments of the zero-mode form factors are
\[
  |k_\ell|=c_\ell Q,
  \qquad c_\ell>0,\qquad \ell=1,\ldots,m,
\]
with \(Q\) much larger than the infrared scale of the theory.  For
differentiated external fermion slots, \(m=4\) and the
\(\mathcal F_\ell\) below are the individual amputated zero-mode form factors
of Proposition~\ref{prop:instanton-individual-zero-mode-form-factor}.  In the
singular-gauge normalization,
\(\mathcal F_\ell(c_\ell Q\rho)=F_{\rm zm}(c_\ell Q\rho/2)\), up to the
fixed spin-color tensor already included in \(\mathcal C_{\rm hard}\).
For a fused bilinear source, \(m=2\) and
\(\mathcal F_\ell(z)=zK_1(z)\).  Keep only the one-loop universal density, the
total four-zero-mode power \(\rho^6\), and the dimensionless hard form-factor
product \(\prod_{\ell=1}^m\mathcal F_\ell(c_\ell Q\rho)\).  Choosing
\(\mu=Q\), the corresponding size factor has the form
\begin{equation}
  \mathcal I_{\rm hard}(Q)
  =
  \exp\!\left[-\frac{8\pi^2}{g_{\rm ht}^2(Q)}+\ii\theta\right]
  \left(\frac{8\pi^2}{g_{\rm ht}^2(Q)}\right)^{2N_c}
  Q^{b_0}
  \int_0^\infty
  \rho^{\,b_0+1}
  \prod_{\ell=1}^{m}
  \mathcal F_\ell(c_\ell Q\rho)\,\dd\rho
  \times\mathcal C_{\rm hard},
  \label{eq:hard-momentum-instanton-size-factor}
\end{equation}
up to finite determinant remainders, orientation tensors, source
normalizations, and higher-loop matching factors.  With \(s=Q\rho\),
\begin{equation}
  Q^{b_0}
  \int_0^\infty
  \rho^{\,b_0+1}
  \prod_{\ell=1}^{m}
  \mathcal F_\ell(c_\ell Q\rho)\,\dd\rho
  =
  Q^{-2}
  \int_0^\infty
  s^{\,b_0+1}
  \prod_{\ell=1}^{m}
  \mathcal F_\ell(c_\ell s)\,\dd s .
  \label{eq:hard-momentum-instanton-size-scaling}
\end{equation}
This \(Q^{-2}\) is the engineering dimension of the source-amputated
four-fermion coefficient after the renormalization point has been set to
\(\mu=Q\).  If the one-loop running exponential is instead traded for the
RG-invariant scale
\[
  \Lambda_{\rm ht}^{b_0}
  =
  Q^{b_0}
  \exp\!\left[-\frac{8\pi^2}{g_{\rm ht}^2(Q)}\right]
  \quad\hbox{at one-loop order},
\]
then the same hard factor is
\begin{equation}
  \mathcal I_{\rm hard}(Q)
  =
  \ee^{\ii\theta}
  \left(\frac{8\pi^2}{g_{\rm ht}^2(Q)}\right)^{2N_c}
  \Lambda_{\rm ht}^{b_0}\,
  Q^{-b_0-2}\,
  \mathcal J_{b_0}(\mathbf c;\mathcal F)\,
  \mathcal C_{\rm hard},
  \label{eq:hard-momentum-instanton-rg-invariant-scaling}
\end{equation}
with
\[
  \mathcal J_{b_0}(\mathbf c;\mathcal F)
  =
  \int_0^\infty
  s^{\,b_0+1}
  \prod_{\ell=1}^{m}
  \mathcal F_\ell(c_\ell s)\,\dd s .
\]
Thus the semiclassical exponential and the size measure together become a
power-suppressed hard coefficient.  The remaining displayed
\((8\pi^2/g_{\rm ht}^2(Q))^{2N_c}\) is the familiar one-loop collective-mode
Jacobian prefactor; higher-loop matching changes the logarithmic factor and
scheme coordinate, not the endpoint mechanism described here.
The small-\(\rho\) endpoint is finite in asymptotically free QCD because the
integrand behaves as \(\rho^{b_0+1}\).  The large-\(\rho\) endpoint depends on
which source construction is being used.  For fused bilinear density entries,
\(zK_1(z)\) gives the exponential cutoff
\[
  \prod_{\ell=1}^{m}\mathcal F_\ell(c_\ell Q\rho)
  =
  O\!\left((Q\rho)^\gamma\exp[-d(\mathbf c)Q\rho]\right),
  \qquad d(\mathbf c)>0,
  \qquad Q\rho\to\infty .
\]
For differentiated singular-gauge fermion slots, however,
\eqref{eq:instanton-individual-zero-mode-form-factor-large} gives only the
power behavior
\[
  \prod_{\ell=1}^{m}F_{\rm zm}(c_\ell Q\rho/2)
  =
  O\!\left((Q\rho)^{-3m}\right)
  \qquad(c_\ell>0).
\]
The large-\(\rho\) part of the size integral is then controlled only if
\begin{equation}
  b_0+1-3m<-1 .
  \label{eq:hard-momentum-individual-slot-large-rho-test}
\end{equation}
For the \(N_f=2\) four-slot hard kernel, \(m=4\), so this condition is
\(b_0<10\); in \(SU(3)\) QCD with two Dirac fundamentals,
\(b_0=29/3\), and the margin is \(1/3\).  If a determinant monomial contains a
soft slot with \(c_\ell=0\), or if the power test fails, the hard source
profile has not by itself produced a controlled large-size instanton
amplitude.  This is a physical endpoint condition in the amplitude integral,
not a cutoff on the instanton moduli space.  The approximation is controlled
only when the dominant range also has \(1/\rho\) in the weak-coupling regime.
\end{controlledapproximation}

\begin{controlledapproximation}[Hard-size dominance versus endpoint convergence]
\label{ca:hard-size-dominance-vs-convergence}
The convergence test in
\eqref{eq:hard-momentum-individual-slot-large-rho-test} is not yet a
dominance statement.  In the notation of
\eqref{eq:hard-momentum-instanton-size-factor}, set
\[
  \mathcal F_{\rm hard}(s)
  =
  \prod_{\ell=1}^{m}\mathcal F_\ell(c_\ell s),
  \qquad
  a=b_0+1 .
\]
After the change of variables \(s=Q\rho\), the absolute tail of the one-loop
hard size integral outside the band \(\rho\le R/Q\) is
\begin{equation}
  Q^{b_0}
  \int_{R/Q}^{\infty}
  \rho^{\,b_0+1}\,
  |\mathcal F_{\rm hard}(Q\rho)|\,\dd\rho
  =
  Q^{-2}
  \int_{R}^{\infty}
  s^{\,a}\,|\mathcal F_{\rm hard}(s)|\,\dd s .
  \label{eq:hard-size-tail-change-variable}
\end{equation}
Thus a statement that the amplitude is dominated by instantons of size
\(\rho\sim Q^{-1}\) requires a bound on the right-hand tail in addition to
the formal local vertex.

For example, suppose that for \(s\ge s_0\)
\[
  |\mathcal F_{\rm hard}(s)|\le C_\infty s^{-\sigma},
  \qquad \sigma>b_0+2 .
\]
Then for \(R\ge s_0\),
\begin{equation}
  Q^{b_0}
  \int_{R/Q}^{\infty}
  \rho^{\,b_0+1}\,
  |\mathcal F_{\rm hard}(Q\rho)|\,\dd\rho
  \le
  Q^{-2}\,
  \frac{C_\infty\,R^{\,b_0+2-\sigma}}
       {\sigma-b_0-2}.
  \label{eq:hard-size-power-tail-bound}
\end{equation}
If instead the fused bilinear-density source gives
\(|\mathcal F_{\rm hard}(s)|\le C_\infty s^\gamma e^{-ds}\), \(d>0\), then
the corresponding tail is bounded by
\[
  Q^{-2}C_\infty
  \int_R^\infty s^{\,b_0+1+\gamma}e^{-ds}\,\dd s ,
\]
and is exponentially small once \(R\) is large.  The two source
constructions therefore have different endpoint control.

For the differentiated singular-gauge four-slot \(SU(3)\), \(N_f=2\) hard
kernel, the individual zero-mode form factors have
\(\sigma=3m=12\), while \(b_0=29/3\).  Hence
\[
  b_0+2-\sigma=-\frac13 .
\]
The contribution from \(\rho>R/Q\) is only \(O(R^{-1/3})\), up to the
spin-color and \(c_\ell\)-dependent constants.  The size integral is
convergent, but the convergence margin is narrow: claiming that the
one-instanton hard kernel is numerically dominated by \(\rho\sim Q^{-1}\)
requires an additional estimate showing that this slowly decaying tail is
small in the chosen source and kinematic window.  This is a physical
endpoint-control hypothesis on the amplitude, not a property of the BPST
moduli coordinates alone.

The same statement is often clearer per logarithmic size shell.  Define the
absolute log-shell density after the hard change of variables by
\begin{equation}
  \mathfrak s_{\rm hard}(s)
  =
  s^{\,b_0+2}\,
  |\mathcal F_{\rm hard}(s)|,
  \qquad
  Q^{b_0}\rho^{\,b_0+1}
  |\mathcal F_{\rm hard}(Q\rho)|\,\dd\rho
  =
  Q^{-2}\mathfrak s_{\rm hard}(s)\,\dd\log s .
  \label{eq:hard-instanton-log-size-shell}
\end{equation}
If \(|\mathcal F_{\rm hard}(s)|\le C_\infty s^{-\sigma}\), the shell and the
tail both decay with the exponent
\[
  \Delta=\sigma-b_0-2 .
\]
For the \(SU(3)\), \(N_f=2\) differentiated four-slot kernel,
\(\Delta=1/3\).  One extra decade in the cutoff \(R\) then suppresses the
tail by only \(10^{-1/3}\), and a tenfold suppression of this power-tail
majorant requires three decades in \(R\).  With the normalized bound used in
\eqref{eq:hard-size-power-tail-bound}, the tail estimate is
\[
  3R^{-1/3};
\]
making this smaller than \(0.1\) requires \(R>27000\).  This number is not a
prediction of the amplitude.  It is a warning about evidence: a formally hard
one-instanton coefficient is not physically dominated by small instantons
until the chosen source profile, kinematics, or matching calculation improves
this log-shell bound.
\end{controlledapproximation}

\begin{controlledapproximation}[Hard source envelope with physical screening]
\label{ca:hard-screened-instanton-size-shell}
The hard and screened size estimates above address different failures of the
same amplitude integral.  A hard source can suppress large instantons through
external form factors, while a physical Higgs, thermal, or medium scale can
screen them through the fluctuation determinant.  When both mechanisms are
present, the dominant size is not obtained by choosing whichever formula is
more convenient.  It is fixed by the combined size density.

Assume that in the selected one-instanton channel the absolute value of the
fully assembled density is bounded in the retained window by
\begin{equation}
  |\mathcal K(\rho)|
  \le
  K_0\,\rho^{A-1}
  \exp[-dQ\rho-m_{\rm scr}^2\rho^2],
  \qquad
  A>0,\quad d>0,\quad m_{\rm scr}>0.
  \label{eq:hard-screened-size-majorant}
\end{equation}
Here \(A=b_0+\beta_{\mathcal X}-4\) contains the collective density,
determinant convention, and zero-mode/source power of the chosen amplitude
channel.  The coefficient \(dQ\) is a proven exponential envelope for the
hard form factors or source wave packets; it is absent for the differentiated
individual zero-mode slots unless such a stronger source bound has actually
been supplied.  The scale \(m_{\rm scr}\) is a physical screening mass from
the same regulated problem, not a hand-inserted cutoff.

The absolute logarithmic size shell of the majorant is
\[
  \rho^A\exp[-dQ\rho-m_{\rm scr}^2\rho^2]\,\dd\log\rho .
\]
Its stationary point solves
\begin{equation}
  2m_{\rm scr}^2\rho_*^2+dQ\rho_*-A=0,
  \qquad
  \rho_*
  =
  \frac{2A}
       {dQ+\sqrt{d^2Q^2+8A m_{\rm scr}^2}}.
  \label{eq:hard-screened-log-shell}
\end{equation}
Thus the hard-only shell \(A/(dQ)\) and the screened-only shell
\(\sqrt{A/2}/m_{\rm scr}\) are limiting coordinates of one combined
problem:
\[
  \frac{\rho_*}{A/(dQ)}
  =
  \frac{2}{1+\sqrt{1+8A m_{\rm scr}^2/(d^2Q^2)}}.
\]
The two limiting diagnostics are
\[
  \epsilon_{\rm hard}=\frac{A m_{\rm scr}}{dQ},
  \qquad
  \epsilon_{\rm scr}=\frac{dQ}{m_{\rm scr}}\sqrt{\frac A2}.
\]
If \(\epsilon_{\rm hard}\ll1\), the hard envelope sets the shell and
screening is a perturbation.  If \(\epsilon_{\rm scr}\ll1\), the physical
screening scale sets the shell and the hard source is only a slowly varying
factor.  In the intermediate regime the two suppressions cooperate and the
dominant size is smaller than either pure estimate.

This criterion is deliberately stronger than convergence.  A hard instanton
coefficient is a controlled short-distance object only if the combined shell
and a surrounding tail bound lie inside a weak-coupling region,
\[
  \Lambda_{\rm QCD}\rho_*\ll1
\]
or inside the corresponding Higgs/thermal weak-coupling window.  If
\(\rho_*\) is instead set by a soft screening scale comparable to the
confinement scale, the formula is at best an absolute majorant on a
semiclassical coordinate; it is not a derivation of a physical QCD amplitude.
This is exactly the point hidden by a moduli-space-only discussion: the
collective coordinate supplies the integration variable, but the determinant,
zero-mode form factors, physical screening, and endpoint estimate decide
whether the instanton sector contributes a controlled observable coefficient.
\end{controlledapproximation}

\begin{controlledapproximation}[\(SU(3)\), two-flavor hard instanton
coefficient]
\label{ca:su3-two-flavor-hard-instanton-coefficient}
Specialize the source-amputated four-slot construction to vectorlike
\(SU(3)\) QCD with \(N_f=2\) Dirac fundamentals.  Use the Pauli--Villars
pure-gauge convention of
Proposition~\ref{prop:pure-gauge-pv-instanton-determinant-constant}, choose a
finite normalization of the light-fermion nonzero-mode determinant in the same
scheme, and suppose all four differentiated external zero-mode slots are hard:
\[
  |k_\ell|=c_\ell Q,\qquad c_\ell>0,\qquad \ell=1,\ldots,4 .
\]
Let \(\mathcal P_{\rm orient}\) denote the shared \(SU(2)\) orientation
projector of
\eqref{eq:instanton-four-slot-orientation-projector}, contracted with the
chosen spin-color source covectors.  In the leading one-instanton hard window,
the short coefficient multiplying the selected four-fermion source tensor has
the form
\begin{equation}
\begin{aligned}
  C^{(4)}_{\rm inst}(Q;R)
  &=
  \ee^{\ii\theta}\,
  C^{\rm PV}_{3}\,\mathcal R_{\rm f}^{\rm PV}(Q)\,
  \left(\frac{8\pi^2}{g_{\rm ht}^2(Q)}\right)^6
  \Lambda_{\rm ht}^{29/3}\,Q^{-35/3}        \\
  &\qquad{}\times
  \mathcal P_{\rm orient}
  \int_0^{R}
    s^{32/3}
    \prod_{\ell=1}^{4}F_{\rm zm}(c_\ell s/2)\,\dd s
  +R_{\rm det}+R_{\rm zm}+R_{\rm src}+R_{\rm Schur}+R_{s>R}.
\end{aligned}
  \label{eq:su3-nf2-hard-instanton-coefficient}
\end{equation}
Here \(C^{\rm PV}_{3}\simeq1.51\times10^{-3}\) is the orientation-integrated
pure-gauge constant in
\eqref{eq:pure-gauge-pv-instanton-density}, while
\(\mathcal R_{\rm f}^{\rm PV}\) is not fixed by the pure-gauge determinant
calculation; it is the finite light-fermion nonzero-mode determinant factor of
the same regulator and source normalization, with the finite-frame covariance
fixed by
Proposition~\ref{prop:instanton-light-fermion-source-frame-covariance}.  The
five remainders are the
determinant, zero-mode/source, source-matching, Schur, and endpoint residuals
of Proposition~\ref{prop:instanton-amplitude-assembly-error-budget}.

The powers in \eqref{eq:su3-nf2-hard-instanton-coefficient} are fixed as
follows.  For \(SU(3)\) with two Dirac fundamentals,
\[
  b_0=\frac{11}{3}3-\frac23\,2=\frac{29}{3}.
\]
The four differentiated zero-mode slots supply \(\rho^6\), so the size
integrand before form factors is \(\rho^{b_0+1}\).  After \(s=Q\rho\) and the
one-loop trade
\(\exp[-8\pi^2/g_{\rm ht}^2(Q)]Q^{b_0}=\Lambda_{\rm ht}^{b_0}\), the hard
coefficient carries
\[
  \Lambda_{\rm ht}^{29/3}Q^{-29/3-2}
  =
  \Lambda_{\rm ht}^{29/3}Q^{-35/3}.
\]
Thus the dimension of the coefficient is \(-2\), as required for a
source-amputated four-fermion operator.  This dimension count is not a
standalone amplitude: the orientation projector, the light-fermion determinant,
the zero-mode form factors, and the endpoint residual are still part of the
calculation.

The endpoint term is especially restrictive.  From
\eqref{eq:instanton-individual-zero-mode-form-factor-large},
\[
  F_{\rm zm}(c_\ell s/2)
  =
  6c_\ell^{-3}s^{-3}+O(s^{-5}),
  \qquad s\to\infty .
\]
Therefore
\begin{equation}
\begin{aligned}
  \int_R^\infty
  s^{32/3}\prod_{\ell=1}^{4}F_{\rm zm}(c_\ell s/2)\,\dd s
  &=
  3\cdot6^4
  \left(\prod_{\ell=1}^{4}c_\ell^{-3}\right)R^{-1/3}
  +O(R^{-7/3}).
\end{aligned}
  \label{eq:su3-nf2-hard-tail-asymptotic}
\end{equation}
So even this hard four-slot coefficient has only a slow \(R^{-1/3}\) endpoint
tail unless the external source construction supplies stronger suppression.
This is the amplitude-level warning in 't Hooft's calculation: the BPST
moduli measure, the determinant constant, and the zero-mode tensor do not yet
give a controlled physical four-point contribution until the hard-source tail
and the remaining remainders have been bounded in the same scheme.
\end{controlledapproximation}

\begin{controlledapproximation}[Finite-window bound for a hard
instanton coefficient]
\label{ca:hard-instanton-finite-window-bound}
The hard coefficient
\eqref{eq:su3-nf2-hard-instanton-coefficient} should be quoted only with a
window error bound.  After the center delta function, source tensor, and
orientation projection have been fixed, write the retained dimensionless
leading size integrand as
\[
  H_{\rm lead}(s)
  =
  s^{32/3}\,
  \mathcal P_{\rm orient}(s)
  \prod_{\ell=1}^{4}F_{\rm zm}(c_\ell s/2),
\]
where \(\mathcal P_{\rm orient}(s)\) includes any source or orientation
weight which remains inside the size integral.  For a cutoff \(R\), define
\begin{equation}
  J_R=\int_0^R H_{\rm lead}(s)\,\dd s,\qquad
  M_R=\int_0^R |H_{\rm lead}(s)|\,\dd s,\qquad
  T_R=\int_R^\infty |H_{\rm lead}(s)|\,\dd s .
  \label{eq:hard-instanton-window-jmt}
\end{equation}
The number \(M_R\) is not the coefficient; it is the absolute mass of the
retained window.  Its role is to measure how much cancellation is present in
the actual signed or complex coefficient \(J_R\).

Suppose the finite-regulator density on \(0<s<R\) differs from the displayed
leading density by multiplicative determinant, zero-mode, and source errors
with absolute window bounds
\[
  |\varepsilon_{\rm det}(s)|\le\epsilon_{\rm det},\qquad
  |\varepsilon_{\rm zm}(s)|\le\epsilon_{\rm zm},\qquad
  |\varepsilon_{\rm src}(s)|\le\epsilon_{\rm src},
\]
and suppose the Schur-complement remainder satisfies
\[
  \int_0^R|\Delta_{\rm Schur}(s)|\,\dd s
  \le\epsilon_{\rm Schur}M_R .
\]
Finally suppose the exact contribution from the omitted tail, including the
difference between the exact and leading kernels beyond \(R\), is bounded by
\[
  |R_{s>R}|\le \tau_R M_R .
\]
Then, after restoring the common prefactor \(\mathcal N(Q)\) in
\eqref{eq:su3-nf2-hard-instanton-coefficient},
\begin{equation}
  \left|
  C_{\rm inst}^{(4),{\rm ex}}(Q)
  -
  \mathcal N(Q)J_R
  \right|
  \le
  |\mathcal N(Q)|\,M_R\,
  \bigl(
    \epsilon_{\rm det}
    +\epsilon_{\rm zm}
    +\epsilon_{\rm src}
    +\epsilon_{\rm Schur}
    +\tau_R
  \bigr).
  \label{eq:hard-instanton-window-absolute-bound}
\end{equation}
If the retained leading window is nondegenerate in the sense that
\[
  |J_R|\ge \kappa_R M_R,\qquad \kappa_R>0,
\]
the same data give the relative error bound
\begin{equation}
  \frac{
  \left|
  C_{\rm inst}^{(4),{\rm ex}}(Q)
  -
  \mathcal N(Q)J_R
  \right|}
  {|\mathcal N(Q)J_R|}
  \le
  \frac{
    \epsilon_{\rm det}
    +\epsilon_{\rm zm}
    +\epsilon_{\rm src}
    +\epsilon_{\rm Schur}
    +\tau_R}
  {\kappa_R}.
  \label{eq:hard-instanton-window-relative-bound}
\end{equation}
Thus endpoint convergence is only one ingredient.  A hard instanton
coefficient can be numerically unstable if the retained window is small after
signed source or orientation cancellations, even when \(T_R/M_R\) is small.

For the differentiated \(SU(3)\), \(N_f=2\) four-slot kernel, a power-tail
majorant of the form
\[
  |H_{\rm lead}(s)|
  \le
  B_\infty s^{-4/3},
  \qquad s\ge R_0,
\]
gives
\begin{equation}
  T_R
  \le
  3B_\infty R^{-1/3},
  \qquad R\ge R_0 .
  \label{eq:hard-instanton-window-tail-bound}
\end{equation}
With only the individual zero-mode large-\(s\) envelope one may take
\(B_\infty\) proportional to
\(6^4\prod_\ell c_\ell^{-3}\) times the source-orientation norm bound.  This
is the quantitative meaning of the slow-tail warning above: a finite hard
coefficient becomes a controlled short-distance observable only after the
chosen source supplies a usable \(M_R\), a noncancellation margin
\(\kappa_R\), and a tail/error budget in the same normalization.  Otherwise
the calculation has produced a valuable source coordinate, not yet a
controlled physical coefficient.
\end{controlledapproximation}

\begin{controlledapproximation}[Four-source instanton amplitude rank-and-error bound]
\label{ca:four-source-instanton-amplitude-rank-bound}
The finite-window coefficient above still assumes that the four external
fermion slots have already been converted into a nondegenerate zero-mode
source tensor.  This is the step which turns the collective-coordinate
integral into the kind of four-point amplitude computed in the original
instanton calculation.  For \(N_f=2\), let
\(h^R_1,h^R_2,h^L_1,h^L_2\) be four fixed external test spinors or amputated
source covectors in the chosen regulator and source frame.  On a selected
instanton chart define the right and left zero-mode overlap matrices
\[
  B_R(z)_{if}=\langle h_i^R,\psi_{0,f}^R(A_z)\rangle,\qquad
  B_L(z)_{jf}=\langle h_j^L,\psi_{0,f}^L(A_z)\rangle,
  \qquad i,j,f=1,2 .
\]
After orientation and spin-color contraction, the fermion zero-mode part of
the four-source instanton integrand contains
\[
  Z_0(z)=\det B_R(z)\det B_L(z).
\]
Thus a hard momentum scale by itself does not guarantee a nonzero four-point
instanton amplitude.  The two right slots and the two left slots must each
span the corresponding zero-mode coefficient plane after the same
orientation and source projection have been applied.

Let \(B_R^0(s,U)\) and \(B_L^0(s,U)\) be the retained leading overlap matrices
inside \(0<s<R\).  Suppose their entries are bounded by \(L_R,L_L\), while
the exact overlaps obey
\[
  |(B_R-B_R^0)_{if}|\le\eta_R,\qquad
  |(B_L-B_L^0)_{jf}|\le\eta_L .
\]
For any \(2\times2\) matrix \(B^0\) with \(|B^0_{ab}|\le L\),
\begin{equation}
  |\det(B^0+\Delta B)-\det B^0|
  \le
  4L\eta+2\eta^2,
  \qquad |\Delta B_{ab}|\le\eta .
  \label{eq:four-source-determinant-stability}
\end{equation}
Consequently, if
\[
  |\det B_R^0|\ge d_R>0,\qquad
  |\det B_L^0|\ge d_L>0
\]
on the retained window, the zero-mode source factor has relative error
\begin{equation}
  \epsilon_{\rm rank}
  \le
  \frac{E_R}{d_R}
  +
  \frac{E_L}{d_L}
  +
  \frac{E_RE_L}{d_Rd_L},
  \qquad
  E_R=4L_R\eta_R+2\eta_R^2,\quad
  E_L=4L_L\eta_L+2\eta_L^2 .
  \label{eq:four-source-rank-error}
\end{equation}
The word ``rank'' in \(\epsilon_{\rm rank}\) is intentional: if either
determinant margin \(d_R\) or \(d_L\) is absent, the four-source amplitude is
not controlled by the BPST measure, by the nonzero-mode determinant, or by the
large-momentum tail estimate.

Write \(J_R\) and \(M_R\) for the signed retained integral and absolute mass
from \eqref{eq:hard-instanton-window-jmt}, now including
\(\det B_R^0\det B_L^0\) in the leading source tensor.  If the determinant,
Schur, projection/matching, and endpoint bounds on the same window are
\(\epsilon_{\rm det}\), \(\epsilon_{\rm Schur}\),
\(\epsilon_{\rm proj}\), and \(\tau_R\), then the Euclidean four-source
coefficient obeys
\begin{equation}
  \left|
    C_{\rm inst}^{(4),{\rm ex}}[\mathfrak h]
    -\mathcal N(Q)J_R[\mathfrak h]
  \right|
  \le
  |\mathcal N(Q)|M_R[\mathfrak h]\,
  \bigl(
    \epsilon_{\rm det}
    +\epsilon_{\rm rank}
    +\epsilon_{\rm Schur}
    +\epsilon_{\rm proj}
    +\tau_R
  \bigr).
  \label{eq:four-source-instanton-amplitude-bound}
\end{equation}
A physical scattering, spectral, or OPE coefficient still requires the
continuation/projection residuals already displayed in
\eqref{eq:instanton-amplitude-pipeline-physical-map}.  The point of
\eqref{eq:four-source-instanton-amplitude-bound} is narrower but crucial:
the hard four-point instanton calculation is not an analysis of the instanton
moduli space alone.  It is a stability statement for a product of the
collective-coordinate density, the fluctuation determinant, the four
zero-mode source overlaps, the size-window tail, and the final observable
projection.
\end{controlledapproximation}

\begin{controlledapproximation}[Instanton size-factor split]
\label{ca:instanton-wilsonian-size-factorization}
Let \(K_{\Lambda}(\rho;\mu,\mathcal J)\) denote the fully paired
finite-regulator one-instanton size integrand for a selected source family:
the center phase, orientation tensor, nonzero-mode determinant, zero-mode
source coefficient, and any anomaly-line pairing have already been included.
Fix a finite upper regulator \(\rho_{\max}\) and a factorization scale
\(\mu_I\) with
\[
  0<\rho_I:=\mu_I^{-1}<\rho_{\max}.
\]
Then the exact finite-regulator contribution on this size window has the
split
\begin{equation}
  \mathcal A_{\Lambda}^{[0,\rho_{\max}]}
  =
  \mathcal A_{\Lambda}^{<}(\mu_I)
  +
  \mathcal A_{\Lambda}^{>}(\mu_I),
  \qquad
  \mathcal A_{\Lambda}^{<}(\mu_I)
  =
  \int_0^{\rho_I}K_\Lambda(\rho)\,\dd\rho ,
  \quad
  \mathcal A_{\Lambda}^{>}(\mu_I)
  =
  \int_{\rho_I}^{\rho_{\max}}K_\Lambda(\rho)\,\dd\rho .
  \label{eq:instanton-size-wilsonian-split}
\end{equation}
If \(K_\Lambda\) is continuous at \(\rho_I\), the factorization-scale
derivatives are the boundary flux through the artificial cut:
\begin{equation}
  \partial_{\log\mu_I}\mathcal A_{\Lambda}^{<}(\mu_I)
  =
  -\rho_I K_\Lambda(\rho_I),
  \qquad
  \partial_{\log\mu_I}\mathcal A_{\Lambda}^{>}(\mu_I)
  =
  +\rho_I K_\Lambda(\rho_I).
  \label{eq:instanton-size-factorization-boundary-flow}
\end{equation}
Thus \(\mathcal A_{\Lambda}^{[0,\rho_{\max}]}\) is independent of
\(\mu_I\), while the short-size piece alone is not.

The semiclassical local 't Hooft vertex is therefore a Wilsonian object.  If
one replaces \(K_\Lambda\) by the one-loop small-size approximation
\(K_{\rm sc}^{<}\) only for \(0<\rho<\rho_I\), then
\[
  C_{\rm inst}^{<}(\mu_I;\mathcal J)
  =
  \int_0^{\rho_I}K_{\rm sc}^{<}(\rho;\mu,\mathcal J)\,\dd\rho
\]
is a short-instanton coefficient.  Its
\(-\rho_IK_{\rm sc}^{<}(\rho_I)\) scale derivative is cancelled only by the
long-distance matrix element or remainder assigned to \(\rho>\rho_I\), up to
the matching error between \(K_\Lambda\) and \(K_{\rm sc}^{<}\) near the cut.
Discarding that long part leaves a spurious \(\mu_I\)-dependence whose size is
precisely the boundary density.  For a hard external scale \(Q\), choosing
\(\rho_I=R/Q\) identifies this boundary term with the same large-size tail
diagnostic in \eqref{eq:hard-size-tail-change-variable}.  The computed
short-instanton coefficient is useful only when this boundary flux and the
remaining \(\rho>\rho_I\) contribution are controlled by the physical source,
screening scale, phase, or matching calculation under consideration.
\end{controlledapproximation}

\begin{controlledapproximation}[Short-instanton OPE coefficient]
\label{ca:short-instanton-ope-coefficient}
The short side of
ControlledApproximation~\ref{ca:instanton-wilsonian-size-factorization} becomes
a local 't Hooft vertex only after it is projected onto a renormalized local
operator basis.  Let \(\{[O_I]_\mu\}_{I\in\mathcal B}\) be a finite
source-normalized basis in the selected quantum-number channel, containing
\([\det(\bar q_Rq_L)]_\mu\) and any operators of the same dimension and
symmetry which are retained by the chosen hard OPE.  Suppose the short-size
part of a gauge-invariant source correlator admits the expansion
\begin{equation}
  G_{\mathcal J,\Lambda}^{<}(\mu_I)
  =
  \sum_{I\in\mathcal B}
    C_I^{<}(\mu_I,\mu)\,
    \left\langle\mathcal J\,[O_I]_\mu\right\rangle
  +
  R_{\rm OPE}^{<}(\mu_I,\mu).
  \label{eq:short-instanton-ope-coefficient-expansion}
\end{equation}
The coefficient \(C_I^{<}\) is obtained by integrating the small-size instanton
density, the nonzero-mode determinant, the zero-mode source coefficient, and
the orientation tensor against the projector onto \([O_I]_\mu\), over
\(0<\rho<\rho_I=\mu_I^{-1}\).  For the leading \(N_f=2\) determinant channel,
this is the coefficient of \([\det(\bar q_Rq_L)]_\mu\); derivative operators,
mass insertions, and source-matching remnants belong either to the retained
basis or to \(R_{\rm OPE}^{<}\).

This projection has two different scale laws.  The renormalization scale
\(\mu\) is the composite-operator scale.  In the convention of
Chapter~\ref{ch:volume-ii-renormalized-operators-minimal-subtraction},
\[
  \frac{\dd}{\dd\log\mu}[O_I]_\mu
  =
  -\gamma_{IK}(\mu)[O_K]_\mu .
\]
Therefore the coefficient row must obey
\begin{equation}
  \frac{\dd}{\dd\log\mu}C_K^{<}(\mu_I,\mu)
  =
  \sum_{I\in\mathcal B}C_I^{<}(\mu_I,\mu)\,\gamma_{IK}(\mu)
  \label{eq:short-instanton-coefficient-rg}
\end{equation}
up to the truncation error \(R_{\rm OPE}^{<}\).  A finite operator-frame change
\([\widetilde O_A]_\mu=M_{AI}(\mu)[O_I]_\mu\) changes the displayed coefficient
by the inverse rule
\begin{equation}
  \widetilde C_A^{<}=C_I^{<}(M^{-1})_{IA},
  \qquad
  \sum_A\widetilde C_A^{<}[\widetilde O_A]_\mu
  =
  \sum_I C_I^{<}[O_I]_\mu .
  \label{eq:short-instanton-coefficient-frame-change}
\end{equation}
Thus the local 't Hooft vertex coefficient is not a scheme-independent number;
only its pairing with the renormalized operator basis and matrix elements is.

The factorization scale \(\mu_I\) is different.  If
\(K_I^{\rm OPE}(\rho;\mu)\) is the \([O_I]_\mu\)-projected short-size
integrand, then
\begin{equation}
  \partial_{\log\mu_I}C_I^{<}(\mu_I,\mu)
  =
  -\rho_I K_I^{\rm OPE}(\rho_I;\mu).
  \label{eq:short-instanton-coefficient-factorization-flow}
\end{equation}
This artificial boundary flow is not cancelled by operator anomalous
dimensions.  It is cancelled by the long-size matrix element or remainder, as
in \eqref{eq:instanton-size-factorization-boundary-flow}.  Consequently a
standalone local 't Hooft operator with no declared \(\mu_I\), no operator
scheme, and no long-distance remainder is not a physical instanton amplitude.
\end{controlledapproximation}

This hard-size window is the amplitude counterpart of the local
't Hooft vertex.  The local operator records the \(Q\rho\ll1\) tensor and
selection rule, while \eqref{eq:hard-momentum-instanton-size-factor} records
how the same zero modes and fluctuation determinant enter an actual hard
correlator.  It also explains why the mass-saturated vacuum activity below is
a different object: a vacuum theta term has no external form factor to
exclude \(\rho\to\infty\), so its large-size behavior is an infrared
question, not a semiclassical prediction.

Thus the \(Q=1\) amplitude is the microscopic origin of the \(U(1)_A\)
selection rule.  The determinant in
\eqref{eq:qcd-thooft-vertex-leading-local} carries axial charge \(2N_f\);
equivalently, a mass-saturated instanton contribution depends on the CP-odd
combination \(\bar\theta=\theta+\arg\det M\) from
\eqref{eq:qcd-physical-theta}.  This gives the same anomalous Jacobian in the
coordinate system where the fermion zero modes of the instanton background
have been integrated.

\begin{controlledapproximation}[Mass-saturated vacuum activity and the
size integral]
\label{ca:mass-saturated-instanton-vacuum-activity}
Assume vectorlike \(SU(N_c)\) QCD with \(N_f\) Dirac fundamentals, a diagonal
renormalized mass matrix \(M(\mu)\), and a one-instanton size window
\(\rho_-<\rho<\rho_+\) in which the one-loop density and the leading
zero-mode mass overlaps are valid.  In the same local normalization as
\eqref{eq:qcd-thooft-vertex-leading-local}, saturating one Dirac zero-mode
pair by the mass term supplies the dimensionless factor \(m_f(\mu)\rho\).
Thus the mass-saturated one-instanton vacuum activity in this window has the
form
\begin{equation}
  \zeta_m^{[\rho_-,\rho_+]}(\mu)
  =
  \mathcal C_{\mathcal S,N_c,N_f}\,
  \left(\frac{8\pi^2}{g_{\rm ht}^2(\mu)}\right)^{2N_c}
  \exp\!\left[-\frac{8\pi^2}{g_{\rm ht}^2(\mu)}\right]
  |\det M(\mu)|\,\mu^{b_0}
  \int_{\rho_-}^{\rho_+}
  \rho^{\,b_0+N_f-5}\,\dd\rho ,
  \label{eq:mass-saturated-instanton-window-activity}
\end{equation}
up to higher-loop/operator-renormalization factors of the chosen scheme.  The
constant \(\mathcal C_{\mathcal S,N_c,N_f}\) contains the pure-gauge
orientation and determinant convention together with the light-fermion
nonzero-mode factor \(\mathcal R_{\rm f}^{\mathcal S}\) of
Proposition~\ref{prop:instanton-fermion-determinant-zero-mode-factorization},
evaluated in the same small-mass window.  It does not include the displayed
\(\prod_f(m_f\rho)\) factor: that product is the zero-mode saturation
determinant.  The corresponding \(Q=1\) vacuum term carries the phase
\begin{equation}
  V_4\zeta_m^{[\rho_-,\rho_+]}
  \exp\!\left[\ii(\theta+\arg\det M)\right],
  \label{eq:mass-saturated-instanton-vacuum-phase}
\end{equation}
while the anti-instanton term carries the complex conjugate phase.

The displayed size integral isolates the two endpoint questions:
\begin{equation}
  \int_{\rho_-}^{\rho_+}
  \rho^{\,b_0+N_f-5}\,\dd\rho
  =
  \begin{cases}
    \dfrac{\rho_+^{\,b_0+N_f-4}-\rho_-^{\,b_0+N_f-4}}
          {b_0+N_f-4},
      & b_0+N_f\ne4,\\[10pt]
    \log(\rho_+/\rho_-),
      & b_0+N_f=4 .
  \end{cases}
  \label{eq:mass-saturated-size-window-integral}
\end{equation}
The small-size endpoint is locally finite precisely when
\(b_0+N_f>4\).  For \(SU(3)\), this margin is
\[
  b_0+N_f-4
  =
  7+\frac{N_f}{3},
\]
so the mass-saturated one-instanton vacuum term is not obstructed at
\(\rho=0\) in the asymptotically free range.  The same positive power means
that extending the one-loop formula to \(\rho_+=\infty\) would give an
infrared power divergence.  That divergence is not a new ultraviolet
counterterm; it says that the zero-temperature large-size region is outside
the controlled weak-coupling instanton calculation.  A finite dilute activity
therefore requires an additional physical infrared cutoff or mechanism, such
as a Higgs scale, a thermal screening scale, or a separately justified
confining/chiral-dynamics input.

If one quark is massless and no external source saturates its zero modes,
\(\det M=0\) and the vacuum activity
\eqref{eq:mass-saturated-instanton-window-activity} vanishes.  This is
compatible with nonzero instanton contributions to fermionic correlators:
those correlators use the source-differentiated zero-mode coefficient in
\eqref{eq:proper-time-four-fermion-instanton-amplitude} instead of the mass
determinant in \eqref{eq:mass-saturated-instanton-window-activity}.
\end{controlledapproximation}

\begin{proposition}[Screened one-instanton size integral]
\label{prop:screened-one-instanton-size-integral}
Let a selected one-instanton amplitude channel have the one-loop small-size
density
\[
  K_{\rm sc}(\rho)
  =
  K_0\,\rho^{A-1},
  \qquad A=b_0+\beta_{\mathcal X}-4>0,
\]
after the collective-coordinate density, nonzero-mode determinant convention,
and zero-mode source coefficient have been assembled.  Suppose a physical
infrared mechanism in the same regulated problem contributes a positive
screening factor
\[
  \exp[-m_{\rm scr}^2\rho^2],
  \qquad m_{\rm scr}>0,
\]
for example from Higgsing, thermal screening, or a declared source envelope.
The screened leading contribution is then
\begin{equation}
  \mathcal A_{\rm scr}^{(0)}
  =
  K_0\int_0^\infty
  \rho^{A-1}\ee^{-m_{\rm scr}^2\rho^2}\,\dd\rho
  =
  \frac{K_0}{2}\,
  m_{\rm scr}^{-A}\,
  \Gamma\!\left(\frac A2\right).
  \label{eq:screened-one-instanton-size-integral}
\end{equation}
In the leading screened density, even moments of the instanton size are
\begin{equation}
  \bigl\langle\rho^{2n}\bigr\rangle_0
  =
  m_{\rm scr}^{-2n}
  \frac{\Gamma(A/2+n)}{\Gamma(A/2)},
  \qquad
  m_{\rm scr}^2\langle\rho^2\rangle_0=\frac A2 .
  \label{eq:screened-one-instanton-size-moments}
\end{equation}
If \(A>1\), the logarithmic size-shell density
\(\rho^{A}\exp[-m_{\rm scr}^2\rho^2]\) is stationary at
\begin{equation}
  m_{\rm scr}^2\rho_{\rm shell}^2=\frac{A}{2},
  \label{eq:screened-one-instanton-log-shell}
\end{equation}
while the ordinary \(\dd\rho\)-density is stationary at
\begin{equation}
  m_{\rm scr}^2\rho_{\rm dens}^2=\frac{A-1}{2}.
  \label{eq:screened-one-instanton-density-saddle}
\end{equation}
For \(SU(3)\) with \(N_f=2\) massive Dirac fundamentals in the mass-saturated
vacuum channel, \(\beta_{\mathcal X}=N_f\) and
\[
  A=b_0+N_f-4=\frac{23}{3}.
\]
Thus the screened leading activity is proportional to
\[
  m_{\rm scr}^{-23/3}\Gamma(23/6),
  \qquad
  m_{\rm scr}^2\langle\rho^2\rangle_0=\frac{23}{6},
  \qquad
  m_{\rm scr}^2\rho_{\rm shell}^2=\frac{23}{6}.
\]
This is a controlled amplitude formula only when the same physical screening
scale keeps the dominant shell inside the weak-coupling and finite-regulator
window.  Sending \(m_{\rm scr}\to0\) recovers the infrared divergence of
\eqref{eq:mass-saturated-size-window-integral}; the screened formula is not a
license to use the one-loop instanton density at arbitrarily large size.
\end{proposition}

\begin{proof}
The substitution \(u=m_{\rm scr}^2\rho^2\) gives
\[
  \int_0^\infty\rho^{A-1}\ee^{-m_{\rm scr}^2\rho^2}\,\dd\rho
  =
  \frac12m_{\rm scr}^{-A}
  \int_0^\infty u^{A/2-1}\ee^{-u}\,\dd u,
\]
which is \eqref{eq:screened-one-instanton-size-integral}.  Dividing the same
integral with \(A\) replaced by \(A+2n\) by the \(n=0\) integral gives
\eqref{eq:screened-one-instanton-size-moments}.  The stationary-shell formulas
come from differentiating
\(A\log\rho-m_{\rm scr}^2\rho^2\) for the \(\dd\log\rho\) shell and
\((A-1)\log\rho-m_{\rm scr}^2\rho^2\) for the ordinary density.  The
\(SU(3)\), \(N_f=2\) specialization substitutes
\(b_0=29/3\) and \(\beta_{\mathcal X}=2\).
\end{proof}

\begin{controlledapproximation}[Thermal determinant screening of the size integral]
\label{ca:thermal-instanton-determinant-screening}
Consider weak-coupling finite-temperature \(SU(N_c)\) QCD with \(N_f\) light
Dirac fundamentals and a one-instanton amplitude channel whose zero-temperature
small-size density has the form used in
Proposition~\ref{prop:screened-one-instanton-size-integral}.  The leading
high-temperature fluctuation determinant around the periodic one-instanton
background supplies the Gaussian size weight
\begin{equation}
  K_T(\rho)
  =
  K_0\,\rho^{A-1}
  \exp[-m_T^2\rho^2]\,
  \exp[R_T(\rho)],
  \qquad
  m_T^2=\pi^2T^2\,\frac{2N_c+N_f}{3}.
  \label{eq:thermal-instanton-screening-density}
\end{equation}
Here \(R_T\) denotes the nonquadratic finite-temperature determinant and
caloron-shape remainder in the same regulator and source convention.  In the
trace-delta coupling convention of
Chapter~\ref{ch:thermal-gauge-theory-screening},
\eqref{eq:one-loop-debye-mass-general} gives
\[
  \frac{m_D^2}{g_{\rm YM}^2}
  =
  T^2\,\frac{2N_c+N_f}{3}
\]
for fundamental Dirac matter, so \(m_T^2=\pi^2m_D^2/g_{\rm YM}^2\).  The
thermal instanton cutoff is therefore a determinant response of the plasma,
not a classical bound on the BPST moduli space.

If the chosen amplitude window controls the residual by
\[
  |R_T(\rho)|\le\varepsilon_T ,
\]
define the Gaussian approximation
\begin{equation}
  \mathcal A_T^{G}
  =
  \frac{K_0}{2}
  \left(\pi^2T^2\frac{2N_c+N_f}{3}\right)^{-A/2}
  \Gamma\!\left(\frac A2\right)
  \label{eq:thermal-instanton-gaussian-amplitude}
\end{equation}
and the absolute Gaussian majorant
\begin{equation}
  \mathcal M_T^{G}
  =
  \frac{|K_0|}{2}
  \left(\pi^2T^2\frac{2N_c+N_f}{3}\right)^{-A/2}
  \Gamma\!\left(\frac A2\right).
  \label{eq:thermal-instanton-gaussian-majorant}
\end{equation}
Since
\(|\ee^{R_T(\rho)}-1|\le\ee^{\varepsilon_T}-1\), the general
source-projected amplitude satisfies the manifestly positive estimate
\begin{equation}
  \left|\mathcal A_T-\mathcal A_T^{G}\right|
  \le
  \bigl(\ee^{\varepsilon_T}-1\bigr)\mathcal M_T^{G}
  \label{eq:thermal-instanton-residual-bound}
\end{equation}
whenever the same size window contains the support of the amplitude.  The
stronger positive-activity form
\[
  \left|\mathcal A_T-\mathcal A_T^{G}\right|
  \le
  \bigl(\ee^{\varepsilon_T}-1\bigr)\mathcal A_T^{G}
\]
is allowed only when \(K_0\ge0\) and the residual is real in the selected
channel, so that the integrand is a positive activity density.  If
\(K_0=\ee^{\ii\varphi}|K_0|\) has a fixed phase and \(R_T\) is real, the same
proof gives the fixed-phase estimate with \(|\mathcal A_T^{G}|\) on the
right-hand side.  For a genuinely complex source amplitude, only
\eqref{eq:thermal-instanton-residual-bound} is an absolute-error bound.  The
leading thermal size moments are the moments in
\eqref{eq:screened-one-instanton-size-moments} with \(m_{\rm scr}=m_T\); in a
complex channel they are moments of the Gaussian majorant, not probability
moments of the amplitude itself.

For the mass-saturated \(SU(3)\), \(N_f=2\) vacuum channel,
\[
  m_T^2=\frac83\pi^2T^2,\qquad
  A=\frac{23}{3},
\]
so
\begin{equation}
  \mathcal A_T^{G}
  =
  \frac{K_0}{2}
  \left(\frac83\pi^2T^2\right)^{-23/6}
  \Gamma\!\left(\frac{23}{6}\right),
  \qquad
  \pi^2T^2\langle\rho^2\rangle_G
  =
  \pi^2T^2\rho_{\rm shell}^2
  =
  \frac{23}{16}.
  \label{eq:thermal-su3-nf2-screened-instanton}
\end{equation}
The formula is useful only in a regime where the periodic-instanton determinant,
the residual \(R_T\), and the source or volume limits are controlled.  Near the
thermal transition, or in zero-temperature QCD after sending \(T\to0\), it
does not replace the large-size dynamics by a semiclassical prediction.
\end{controlledapproximation}

\begin{controlledapproximation}[Dilute instanton gas as an amplitude expansion]
\label{ca:dilute-instanton-gas-amplitude}
Suppose the regulated one-instanton and one-anti-instanton amplitudes obtained
above admit finite, translation-invariant vacuum activities \(\zeta_\Lambda\)
per Euclidean four-volume after the size, orientation, nonzero-mode
determinant, and any required zero-mode mass-overlap factors have all been
included.  Assume further that the multi-instanton cluster expansion has a
window in which overlap corrections, instanton--anti-instanton interactions,
and boundary-face remainders contribute \(o(V_4)\) to \(\log Z\).  Then the
leading dilute-gas theta dependence is not a new moduli-space statement; it is
the Poisson sum of the already constructed amplitudes:
\begin{equation}
  \frac{Z_{\rm dig}(\bar\theta)}{Z_0}
  =
  \sum_{n_+,n_-\ge0}
  \frac{(V_4\zeta_\Lambda)^{n_+}}{n_+!}
  \frac{(V_4\zeta_\Lambda)^{n_-}}{n_-!}
  \ee^{\ii(n_+-n_-)\bar\theta}
  =
  \exp\!\left(2V_4\zeta_\Lambda\cos\bar\theta\right).
  \label{eq:dilute-instanton-gas-partition}
\end{equation}
Thus the vacuum-energy difference in this approximation is
\begin{equation}
  E_{\rm dig}(\bar\theta)-E_{\rm dig}(0)
  =
  2\zeta_\Lambda(1-\cos\bar\theta),
  \label{eq:dilute-instanton-gas-energy}
\end{equation}
and the topological-charge cumulants at \(\bar\theta=0\) obey
\begin{equation}
  \kappa_{2m}(Q)=2V_4\zeta_\Lambda,
  \qquad
  \kappa_{2m+1}(Q)=0,
  \qquad
  \chi_{\rm top}^{\rm dig}=2\zeta_\Lambda .
  \label{eq:dilute-instanton-gas-cumulants}
\end{equation}
Equivalently, writing
\[
  E_{\rm dig}(\bar\theta)-E_{\rm dig}(0)
  =
  \frac12\chi_{\rm top}^{\rm dig}\bar\theta^2
  \left(
    1+b_2^{\rm dig}\bar\theta^2+O(\bar\theta^4)
  \right)
\]
gives \(b_2^{\rm dig}=-1/12\).

This is a controlled semiclassical amplitude expansion only when the activity
is finite and dilute in the relevant correlation volume.  If a massless quark
zero mode is not saturated by a mass or source, the vacuum activity for this
term vanishes; instanton sectors may still contribute to fermionic correlators
such as \eqref{eq:proper-time-four-fermion-instanton-amplitude}.  If the
\(\rho\)-integral is infrared divergent, if small-instanton boundary terms are
not controlled, or if instanton interactions are not negligible, then the
Poisson form \eqref{eq:dilute-instanton-gas-partition} has not been derived.
In particular, this dilute four-dimensional gas is not a derivation of
confinement in zero-temperature Yang--Mills.  Its output is the conditional
theta-dependence and zero-mode-saturated amplitude data above.
\end{controlledapproximation}

\begin{controlledapproximation}[First cluster correction to the dilute
instanton gas]
\label{ca:first-cluster-correction-dilute-instantons}
The Poisson formula
\eqref{eq:dilute-instanton-gas-partition} is the zeroth term in a cluster
expansion.  The collective-coordinate chart supplies the integration
variables, while the finite-regulator amplitude scheme must also supply
finite two-body connected cluster integrals \(\mathfrak B_{\sigma\tau}\),
\(\sigma,\tau\in\{+,-\}\), after all relative centers, sizes, orientations,
nonzero-mode determinants, zero-mode saturations, and interaction remainders
in the selected window have been included.  With
\(z_+=\zeta_\Lambda\ee^{\ii\bar\theta}\) and
\(z_-=\zeta_\Lambda\ee^{-\ii\bar\theta}\), the first Mayer correction to the
logarithm has the local form
\begin{equation}
  \frac1{V_4}\log\frac{Z_{\rm cl}(\bar\theta)}{Z_0}
  =
  z_+ + z_-
  +
  \frac12
  \sum_{\sigma,\tau=\pm}
    z_\sigma z_\tau\,\mathfrak B_{\sigma\tau}
  +O(\zeta_\Lambda^3).
  \label{eq:instanton-first-mayer-pressure}
\end{equation}
Assume CP symmetry in this window, so
\[
  \mathfrak B_{++}=\mathfrak B_{--}=:\mathfrak B_2,
  \qquad
  \mathfrak B_{+-}=\mathfrak B_{-+}=:\mathfrak B_0 .
\]
Then
\begin{equation}
  \frac1{V_4}\log\frac{Z_{\rm cl}(\bar\theta)}{Z_0}
  =
  2\zeta_\Lambda\cos\bar\theta
  +
  \zeta_\Lambda^2
  \bigl(\mathfrak B_2\cos(2\bar\theta)+\mathfrak B_0\bigr)
  +O(\zeta_\Lambda^3).
  \label{eq:instanton-first-mayer-theta-pressure}
\end{equation}
The neutral instanton--anti-instanton molecule \(\mathfrak B_0\) changes the
vacuum pressure and the validity of the dilute expansion, but it cancels from
the energy difference \(E(\bar\theta)-E(0)\) at this order.  The same-charge
clusters carry topological charge \(\pm2\), and therefore generate a second
theta harmonic:
\begin{equation}
  E_{\rm cl}(\bar\theta)-E_{\rm cl}(0)
  =
  2\zeta_\Lambda(1-\cos\bar\theta)
  +
  \zeta_\Lambda^2\mathfrak B_2(1-\cos(2\bar\theta))
  +O(\zeta_\Lambda^3).
  \label{eq:instanton-first-cluster-energy}
\end{equation}
Consequently
\begin{equation}
  \chi_{\rm top}^{\rm cl}
  =
  2\zeta_\Lambda+4\zeta_\Lambda^2\mathfrak B_2+O(\zeta_\Lambda^3),
  \qquad
  \partial_{\bar\theta}^4E_{\rm cl}\big|_{0}
  =
  -2\zeta_\Lambda-16\zeta_\Lambda^2\mathfrak B_2+O(\zeta_\Lambda^3),
  \label{eq:instanton-first-cluster-curvatures}
\end{equation}
and, in the expansion convention of
\eqref{eq:dilute-instanton-gas-energy},
\begin{equation}
  b_2^{\rm cl}
  =
  -\frac{1+8\zeta_\Lambda\mathfrak B_2}
         {12(1+2\zeta_\Lambda\mathfrak B_2)}
  +O(\zeta_\Lambda^2)
  =
  -\frac1{12}-\frac12\zeta_\Lambda\mathfrak B_2
  +O(\zeta_\Lambda^2).
  \label{eq:instanton-first-cluster-b2}
\end{equation}
Thus \(b_2=-1/12\) is not a universal instanton prediction.  It is the
Poisson limit of independent \(Q=\pm1\) amplitudes.  The first same-charge
connected cluster already changes the fourth theta curvature, while a neutral
instanton--anti-instanton molecule first tests the cluster expansion and
neutral observables rather than the theta dependence at this order.

The identities above are only as controlled as the two-body cluster
integrals.  They require a single regulator and source convention for the
one-body amplitudes, relative collective-coordinate measure, fluctuation
determinants, zero-mode overlap factors, and boundary-face estimates.  They
do not derive \(\mathfrak B_{\sigma\tau}\) from ADHM data without the
determinant and interaction estimates, nor do they prove that the omitted
three-body and higher clusters are small.
\end{controlledapproximation}

\begin{controlledapproximation}[Thermal dilute topological susceptibility]
\label{ca:thermal-dilute-instanton-susceptibility}
Combine the thermal determinant-screened one-instanton activity of
ControlledApproximation~\ref{ca:thermal-instanton-determinant-screening} with
the dilute-gas hypothesis in
ControlledApproximation~\ref{ca:dilute-instanton-gas-amplitude}.  Write the
Gaussian one-instanton activity per Euclidean four-volume in a positive
mass/source-saturated channel as
\begin{equation}
  \zeta_T^{G}
  =
  \frac{K_{T,0}}{2}
  \left(\pi^2T^2\,\frac{2N_c+N_f}{3}\right)^{-A/2}
  \Gamma\!\left(\frac A2\right),
  \qquad
  A=b_0+\beta_{\mathcal X}-4,
  \label{eq:thermal-dilute-gaussian-activity}
\end{equation}
where \(K_{T,0}\) contains the collective-coordinate normalization,
zero-temperature nonzero-mode determinant convention, mass/source
zero-mode-saturation coefficient, and any finite source-frame normalization in
the selected channel.  Here \(K_{T,0}\ge0\) and the residual is real; otherwise
one has a source amplitude controlled by the majorant
\eqref{eq:thermal-instanton-gaussian-majorant}, not a positive dilute-gas
activity.  If the finite-temperature determinant residual obeys
\(|R_T(\rho)|\le\varepsilon_T\) on the same size window, the exact positive
activity \(\zeta_T\) satisfies
\begin{equation}
  \left|\zeta_T-\zeta_T^{G}\right|
  \le
  \bigl(\ee^{\varepsilon_T}-1\bigr)\zeta_T^{G}.
  \label{eq:thermal-dilute-activity-residual}
\end{equation}
Consequently the leading dilute topological susceptibility and fourth theta
curvature are
\begin{equation}
  \chi_{\rm top}^{T,{\rm dig}}
  =
  2\zeta_T,\qquad
  \partial_\theta^4E_T^{\rm dig}\big|_{\theta=0}
  =
  -2\zeta_T,\qquad
  b_2^{T,{\rm dig}}=-\frac1{12},
  \label{eq:thermal-dilute-topological-cumulants}
\end{equation}
with Gaussian approximants obtained by replacing \(\zeta_T\) by
\(\zeta_T^G\).  The susceptibility error is bounded by
\begin{equation}
  \left|
  \chi_{\rm top}^{T,{\rm dig}}-\chi_{\rm top}^{T,G}
  \right|
  \le
  \bigl(\ee^{\varepsilon_T}-1\bigr)\chi_{\rm top}^{T,G},
  \qquad
  \chi_{\rm top}^{T,G}=2\zeta_T^{G}.
  \label{eq:thermal-dilute-susceptibility-residual}
\end{equation}

For mass-saturated \(SU(3)\), \(N_f=2\) QCD,
\[
  A=\frac{23}{3},\qquad
  \chi_{\rm top}^{T,G}
  =
  K_{T,0}
  \left(\frac83\pi^2T^2\right)^{-23/6}
  \Gamma\!\left(\frac{23}{6}\right).
\]
After the one-loop action has been traded for
\(\Lambda_{\rm ht}^{29/3}\) in the hard convention used above, this is the
dimension-four scaling
\begin{equation}
  \chi_{\rm top}^{T,G}
  \propto
  |m_u m_d|\,
  \Lambda_{\rm ht}^{29/3}\,
  T^{-23/3}
  \label{eq:thermal-su3-nf2-topological-susceptibility-scaling}
\end{equation}
up to the displayed finite determinant, source-frame, and running-mass
coordinates.  This is a statement about the theta curvature of a dilute,
screened thermal instanton ensemble.  It is not a proof that isolated
instantons dominate the topological susceptibility near the crossover or at
zero temperature.
\end{controlledapproximation}

\begin{controlledapproximation}[Topological diffusion and axial-charge
relaxation]
\label{ca:qcd-topological-diffusion-axial-relaxation}
The thermal instanton susceptibility in
ControlledApproximation~\ref{ca:thermal-dilute-instanton-susceptibility} is a
Euclidean theta-curvature observable.  It is not yet a real-time chirality
relaxation rate.  Let
\[
  Q_5(t)=\int\dd^3x\,J_5^0(t,\mathbf x),\qquad
  n_5=Q_5/V,\qquad
  \mathcal Q(t)=\int\dd^3x\,q(t,\mathbf x),
\]
where \(q=(32\pi^2)^{-1}F^a\widetilde F^a\) in the same topological-charge
normalization used above.  For \(N_f\) Dirac flavors with mass matrix \(M\),
the singlet anomaly gives
\begin{equation}
  \dot Q_5(t)
  =
  2N_f\,\mathcal Q(t)
  +
  \Xi_m(t),
  \qquad
  \Xi_m(t)=2\ii\int\dd^3x\,\bar q M\gamma_5 q(t,\mathbf x).
  \label{eq:qcd-axial-charge-anomaly-time}
\end{equation}
In the massless limit \(\Xi_m=0\), but for massive quarks the chirality-flip
operator is a separate real-time datum.

Let \(\rho_{AB}(\omega)\) be the zero-momentum retarded spectral matrix of the
spatially integrated operators \(A,B\in\{\mathcal Q,\Xi_m\}\), divided by
volume after the thermodynamic limit.  The KMS relation defines the
zero-frequency symmetric weights
\begin{equation}
  \Gamma_{AB}^{(0)}
  =
  \lim_{\omega\downarrow0}
  \frac{T}{\omega}\rho_{AB}(\omega),
  \label{eq:qcd-axial-kms-zero-frequency-weights}
\end{equation}
whenever the limit exists in the selected hydrodynamic window.  The topological
entry is the Chern--Simons diffusion rate,
\[
  \Gamma_{\rm CS}=\Gamma_{qq}^{(0)}
  =
  \lim_{Vt\to\infty}
  \frac{\left\langle
  (N_{\rm CS}(t)-N_{\rm CS}(0))^2\right\rangle}{Vt}.
\]
By contrast, after the local \(\theta\)-contact convention of
Definition~\ref{def:topological-susceptibility} has been fixed, the propagated
part of the Euclidean static susceptibility is the integral
\begin{equation}
  \chi_{\rm top}^{E,{\rm prop}}
  =
  \int_0^\beta\dd\tau\int\dd^3x\,
    \langle q(\tau,\mathbf x)q(0,\mathbf0)\rangle_{\rm c}
  =
  \int_0^\infty\frac{\dd\omega}{\pi}\,
    \frac{2\rho_{qq}(\omega)}{\omega}.
  \label{eq:qcd-euclidean-topological-susceptibility-spectral}
\end{equation}
Equations
\eqref{eq:qcd-axial-kms-zero-frequency-weights} and
\eqref{eq:qcd-euclidean-topological-susceptibility-spectral} therefore ask
different questions: the static theta curvature measures a Euclidean spectral
area, together with the declared local contact convention, while
\(\Gamma_{\rm CS}\) measures the low-frequency slope selected by real-time
diffusion.  A dilute instanton calculation of \(\chi_{\rm top}^{T}\) can
control a static theta curvature without fixing the transport coefficient.

If the long-wavelength axial free energy is
\[
  f_5(n_5)=\frac{n_5^2}{2\chi_5},\qquad
  \mu_5=\frac{n_5}{\chi_5},
\]
then linear response gives
\begin{equation}
  \frac{\dd n_5}{\dd t}
  =
  -
  \frac{\Gamma_{\dot Q_5}^{(0)}}{2\chi_5T}\,n_5,
  \qquad
  \Gamma_{\dot Q_5}^{(0)}
  =
  (2N_f)^2\Gamma_{qq}^{(0)}
  +
  4N_f\Gamma_{qm}^{(0)}
  +
  \Gamma_{mm}^{(0)} .
  \label{eq:qcd-axial-charge-relaxation-rate}
\end{equation}
For massless quarks this reduces to the topological contribution
\begin{equation}
  \gamma_5^{\rm top}
  =
  \frac{(2N_f)^2\Gamma_{\rm CS}}{2\chi_5T}.
  \label{eq:qcd-massless-topological-axial-relaxation}
\end{equation}
The factor \(1/2\) is fixed by the displayed variance convention for
\(\Gamma_{\rm CS}\): diffusion of \(N_{\rm CS}\) has Fokker--Planck
coefficient \(\Gamma_{\rm CS}V/2\), and a step of axial charge
\(2N_f\) then drifts in the free energy \(Vn_5^2/(2\chi_5)\) with coefficient
\((2N_f)^2\Gamma_{\rm CS}/(2\chi_5T)\).
Thus an instanton-induced Euclidean \(U(1)_A\) susceptibility, a theta
curvature, and a real-time axial relaxation rate are related by the anomaly
but are not the same observable.  The rate requires the retarded low-frequency
spectral problem, the axial susceptibility of the thermal plasma, and, when
the quarks are massive, the pseudoscalar mass-term and cross spectral weights.
\end{controlledapproximation}

\begin{proposition}[Zero-mode overlap determinant in an instanton ensemble]
\label{prop:instanton-ensemble-zero-mode-overlap}
Fix a finite-volume, finite-regulator vectorlike \(SU(N_c)\) QCD background
which contains \(n_+\) instanton cores and \(n_-\) anti-instanton cores in a
window where the individual zero modes are identifiable and have been
orthonormalized.  Let \(\psi_I^+\), \(I=1,\ldots,n_+\), denote the isolated
zero-mode basis in the chirality sector carried by the instanton cores and let
\(\psi_{\bar J}^-\), \(\bar J=1,\ldots,n_-\), denote the opposite-chirality
basis carried by the anti-instanton cores.  The leading overlap matrix is
\begin{equation}
  T_{I\bar J}
  =
  \bigl\langle
    \psi_I^+,\mathcal D_E[A_{\rm ens}]\psi_{\bar J}^-
  \bigr\rangle ,
  \qquad
  T:\mathbb C^{n_-}\longrightarrow\mathbb C^{n_+},
  \label{eq:instanton-ensemble-overlap-matrix}
\end{equation}
where \(A_{\rm ens}\) is the chosen multi-lump gauge field.  After the
nonzero-mode determinant and source-convention factors have been separated as
in Proposition~\ref{prop:instanton-fermion-determinant-zero-mode-factorization},
the one-flavor massive Dirac operator restricted to this near-zero-mode
subspace is
\begin{equation}
  \mathcal D_{\rm zm}(m)
  =
  \begin{pmatrix}
    m\,\mathbf 1_{n_+} & T\\
    -T^\dagger & m\,\mathbf 1_{n_-}
  \end{pmatrix}
  +E_{\rm zm},
  \label{eq:instanton-ensemble-zero-mode-block}
\end{equation}
where \(E_{\rm zm}\) collects Gram errors, regulator chirality-breaking terms,
and Schur-complement terms from the nonzero-mode sector.  If
\(s_1,\ldots,s_r\) are the singular values of \(T\),
\(r=\min(n_+,n_-)\), then in the leading projected block
\(E_{\rm zm}=0\),
\begin{equation}
  \det\mathcal D_{\rm zm}(m)
  =
  m^{|n_+-n_-|}
  \prod_{\alpha=1}^{r}\left(m^2+s_\alpha^2\right),
  \label{eq:instanton-ensemble-overlap-determinant}
\end{equation}
and, for real \(m>0\),
\begin{equation}
  \operatorname{Tr}\mathcal D_{\rm zm}(m)^{-1}
  =
  \frac{|n_+-n_-|}{m}
  +
  2m\sum_{\alpha=1}^{r}\frac1{m^2+s_\alpha^2}.
  \label{eq:instanton-ensemble-overlap-resolvent}
\end{equation}
For \(N_f\) Dirac flavors with masses \(m_f\), the zero-mode part of the
leading-block fermion determinant is the product of
\eqref{eq:instanton-ensemble-overlap-determinant} over \(f\).  Hence, inside
the leading projected block, a sector with \(n_+\ne n_-\) has
\(|n_+-n_-|\) unpaired zero modes at \(m=0\), while a balanced
instanton--anti-instanton ensemble can acquire small near-zero Dirac
eigenvalues from the singular values of \(T\).

The corresponding exact statement for the full finite-regulator Dirac
operator requires stronger hypotheses.  If the regulator preserves an exact
chiral/index structure and the full gauge field has topological charge
\(Q=n_+-n_-\), then the full massless operator has
\(|Q|\) index-protected zero modes of the corresponding chirality, up to the
usual convention for whether one counts the kernel of \(\mathcal D\) or
\(\mathcal D^\dagger\).  A generic nonzero \(E_{\rm zm}\) may lift zero modes
of the leading projected block; when \(E_{\rm zm}\) is small, the lifted
near-zero eigenvalues are controlled by the finite-dimensional perturbation
size of \(E_{\rm zm}\), while the index-protected kernel survives only in an
index-preserving setup.
\end{proposition}

\begin{proof}
The massless Euclidean Dirac operator anticommutes with \(\gamma_5\), so in a
chiral zero-mode basis it has only off-diagonal matrix elements at leading
order.  The choice of signs in \eqref{eq:instanton-ensemble-zero-mode-block} is
the anti-Hermitian Euclidean convention; replacing it by an equivalent
Hermitian convention changes phases, not the positive factors below.  Take a
singular-value decomposition \(T=U\Sigma V^\dagger\).  The unitary change of
basis \(\operatorname{diag}(U,V)\) preserves the determinant and decomposes the
block with \(E_{\rm zm}=0\) into \(r\) two-dimensional blocks
\[
  \begin{pmatrix}m&s_\alpha\\ -s_\alpha&m\end{pmatrix}
\]
plus \(|n_+-n_-|\) unpaired one-dimensional \(m\)-blocks.  Each paired block
has determinant \(m^2+s_\alpha^2\) and inverse trace
\(2m/(m^2+s_\alpha^2)\); each unpaired block contributes \(m\) to the
determinant and \(1/m\) to the trace.  Multiplying these factors gives
\eqref{eq:instanton-ensemble-overlap-determinant} and
\eqref{eq:instanton-ensemble-overlap-resolvent}.  Different flavors have
independent Grassmann coefficient spaces, so their finite determinants
multiply.

The exact-kernel statement is a separate index input.  The displayed
singular-value reduction diagonalizes the projected leading block with
\(E_{\rm zm}=0\).  Adding \(E_{\rm zm}\) changes the finite matrix being
diagonalized; ordinary finite-dimensional perturbation estimates bound the
movement of the leading-block eigenvalues by the size of this remainder, but
they do not turn the leading block count into an index theorem.  Exact
full-operator zero modes follow only when the regulator supplies an exact
chiral index and the total charge of the full background is the stated
\(Q=n_+-n_-\).
\end{proof}

\begin{controlledapproximation}[Instanton-liquid spectral-density criterion]
\label{ca:instanton-liquid-spectral-density-criterion}
The determinant \eqref{eq:instanton-ensemble-overlap-determinant} is the finite
linear-algebra core of the instanton-liquid mechanism
\cite{DiakonovPetrov1986InstantonVacuum}.  It is not a statement about the
instanton moduli space by itself.  A physical ensemble measure must also include
the bosonic action, collective-coordinate Jacobians, pure-gauge and fermionic
nonzero-mode determinants, instanton--anti-instanton interactions, regulator
remainders \(E_{\rm zm}\), and the determinant reweighting from the overlap
matrix.  Suppose that, after all those weights are included, the positive
singular-value density
\begin{equation}
  \rho_{\rm zm,\mathcal V_4}(s)
  =
  \frac1{\mathcal V_4}
  \left\langle
    \sum_{\alpha}\delta(s-s_\alpha)
  \right\rangle_{\rm ens}
  \label{eq:instanton-liquid-singular-density}
\end{equation}
has a thermodynamic limit \(\rho_{\rm zm}(s)\) continuous at \(s=0\), and that
the unpaired topological zero-mode density
\(|n_+-n_-|/\mathcal V_4\) vanishes in the same limit.  Then the zero-mode-zone
contribution to the scalar-source derivative has the Banks--Casher form
\begin{equation}
  \Sigma_{\rm zm}
  =
  \lim_{m\downarrow0}
  \int_0^\infty
  \frac{2m\,\rho_{\rm zm}(s)}{m^2+s^2}\,\dd s
  =
  \pi\rho_{\rm zm}(0),
  \label{eq:instanton-liquid-banks-casher-criterion}
\end{equation}
in the normalization of
\eqref{eq:qcd-phase-sigma-definition}--\eqref{eq:qcd-phase-banks-casher}.

This is a criterion, not a derivation of the QCD chiral condensate.  The hard
physics is the generation and control of a finite density of small singular
values of the overlap matrix after the fluctuation determinants and interactions
are included.  A dilute Poisson gas with finite isolated activities gives
theta-cumulants as in \eqref{eq:dilute-instanton-gas-cumulants}; it does not by
itself prove \(\rho_{\rm zm}(0)>0\).  Conversely, if a controlled ensemble has a
nonzero near-zero singular-value density, then the instanton zero modes have
become physical near-zero Dirac eigenstates, and the connection to chiral
symmetry breaking is the spectral statement above rather than another
collective-coordinate volume formula.
\end{controlledapproximation}

\begin{controlledapproximation}[Zero-mode-zone \(U(1)_A\) susceptibility
criterion]
\label{ca:instanton-zero-mode-zone-u1a-susceptibility}
The same leading zero-mode-zone block distinguishes two different physical
questions.  A nonzero \(\rho_{\rm zm}(0)\) gives the condensate criterion
\eqref{eq:instanton-liquid-banks-casher-criterion}; restoration of the
\(U(1)_A\)-paired isovector scalar and pseudoscalar susceptibilities is a
stronger infrared test.  Use the source convention of
\eqref{eq:qcd-axial-current-pseudoscalar-density} and the local curvature
normalization of
\eqref{eq:two-flavor-instanton-source-curvature-functional}--\eqref{eq:two-flavor-instanton-u1a-susceptibility-splitting}.
For a finite leading block with positive singular values \(s_\alpha\), define
the zero-mode-zone spectral contribution to the \(U(1)_A\)-odd splitting by
\begin{equation}
  \Delta_{\pi-\delta}^{\rm zm}(m;\mathcal V_4)
  =
  \frac{2|n_+-n_-|}{\mathcal V_4m^2}
  +
  \frac{4m^2}{\mathcal V_4}
  \sum_{\alpha}
  \frac1{(m^2+s_\alpha^2)^2}.
  \label{eq:instanton-zero-mode-zone-u1a-finite}
\end{equation}
The first term is the exact topological zero-mode pole in this projected
coordinate; it must be removed by taking the thermodynamic limit before the
chiral limit, just as in the Banks--Casher statement.  The paired term follows
directly from the source derivatives of a two-dimensional chiral block:
for \(D_{\alpha,m}=\left(\begin{smallmatrix}m&s_\alpha\\
-s_\alpha&m\end{smallmatrix}\right)\) and
\(\gamma_5=\operatorname{diag}(1,-1)\),
\[
  \operatorname{Tr}(D_{\alpha,m}^{-1}\gamma_5D_{\alpha,m}^{-1}\gamma_5)
  +
  \operatorname{Tr}(D_{\alpha,m}^{-1}D_{\alpha,m}^{-1})
  =
  \frac{4m^2}{(m^2+s_\alpha^2)^2}.
\]
The plus sign is the Euclidean source sign: differentiating twice with respect
to the scalar source gives \(-\operatorname{Tr}D^{-1}D^{-1}\), whereas the
pseudoscalar source \(i\gamma_5p\) gives
\(+\operatorname{Tr}D^{-1}\gamma_5D^{-1}\gamma_5\).

Assume now that the ensemble measure in
\eqref{eq:instanton-liquid-singular-density} has a thermodynamic limit, the
topological zero-mode density vanishes first, and the ultraviolet part of the
near-zero-zone density is locally dominated so that the small-\(s\) behavior
controls the chiral singularity.  Then
\begin{equation}
  \Delta_{\pi-\delta}^{\rm zm}(m)
  =
  4m^2\int_0^\infty
  \frac{\rho_{\rm zm}(s)}{(m^2+s^2)^2}\,\dd s .
  \label{eq:instanton-zero-mode-zone-u1a-kernel}
\end{equation}
Consequently
\begin{align}
  \rho_{\rm zm}(s)=\rho_0+o(1)
  &\quad\Longrightarrow\quad
  \Delta_{\pi-\delta}^{\rm zm}(m)
  =
  \frac{\pi\rho_0}{m}+o(m^{-1}), \\
  \rho_{\rm zm}(s)=c_1s+o(s)
  &\quad\Longrightarrow\quad
  \Delta_{\pi-\delta}^{\rm zm}(m)
  =
  2c_1+o(1), \\
  \rho_{\rm zm}(s)=O(s^{1+\epsilon}),\quad \epsilon>0
  &\quad\Longrightarrow\quad
  \Delta_{\pi-\delta}^{\rm zm}(m)\longrightarrow0
\end{align}
under the same local domination hypothesis.  Thus an instanton ensemble that
produces a Banks--Casher condensate automatically leaves a divergent
\(U(1)_A\)-odd zero-mode-zone susceptibility.  Even an ensemble with
\(\rho_{\rm zm}(0)=0\) can retain a finite \(U(1)_A\)-odd remnant if the
near-zero density is only linear.  Suppression of this channel in the chiral
limit requires a stronger depletion of the overlap singular values than the
vanishing of the condensate alone.

This criterion is deliberately phrased in terms of the weighted ensemble
spectral density, not the geometry of isolated instantons.  The ADHM or
collective-coordinate measure supplies only part of the input.  The physical
question is whether the full fluctuation determinant, fermion determinant,
instanton--anti-instanton interaction, and regulator remainder produce a
specific near-zero singular-value law for \(T\); that law is what decides the
\(\pi\)-\(\delta\) channel.
\end{controlledapproximation}

\begin{controlledapproximation}[Mode-count criterion for the \(U(1)_A\)-odd
channel]
\label{ca:instanton-u1a-mode-count-restoration}
The previous criterion can be stated without assuming that the near-zero
singular values already have a smooth density.  At finite volume define the
positive counting function
\begin{equation}
  N_{\rm zm,\mathcal V_4}(\lambda)
  =
  \frac1{\mathcal V_4}
  \left\langle
    \#\{\alpha:0<s_\alpha\le\lambda\}
  \right\rangle_{\rm ens},
  \qquad \lambda>0,
  \label{eq:instanton-u1a-mode-counting-function}
\end{equation}
and let
\(\nu_{\mathcal V_4}=\mathcal V_4^{-1}\langle |n_+-n_-|\rangle_{\rm ens}\)
be the density of index-protected unpaired modes in the same ensemble
coordinate.  The finite zero-mode-zone splitting may be written as the
Stieltjes integral
\begin{equation}
  \Delta_{\pi-\delta}^{\rm zm}(m;\mathcal V_4)
  =
  \frac{2\nu_{\mathcal V_4}}{m^2}
  +
  4m^2\int_{(0,\infty)}
  \frac{\dd N_{\rm zm,\mathcal V_4}(s)}{(m^2+s^2)^2}.
  \label{eq:instanton-u1a-mode-count-stieltjes}
\end{equation}
Thus the order of limits is part of the observable.  At fixed finite volume
and nonzero index, the first term diverges as \(m^{-2}\).  The thermodynamic
limit must first make \(\nu_{\mathcal V_4}\to0\) before the chiral limit is
used to test the paired near-zero zone.

Fix a small \(\delta>0\).  Suppose that, after the thermodynamic limit and
after removing the index-protected density, the paired singular values obey a
mode-count bound
\begin{equation}
  N_{\rm zm}(\lambda)
  \le
  C\,\lambda^{2+\varepsilon},
  \qquad 0<\lambda\le\delta,
  \qquad \varepsilon>0,
  \label{eq:instanton-u1a-superlinear-mode-count-bound}
\end{equation}
and that the total retained near-zero-zone density above \(\delta\) is
bounded.  Split the Stieltjes integral in
\eqref{eq:instanton-u1a-mode-count-stieltjes} into
\((0,m]\), dyadic shells \((2^jm,2^{j+1}m]\) below \(\delta\), and the
remaining interval above \(\delta\).  The first piece is at most
\[
  \frac{4N_{\rm zm}(m)}{m^2}
  \le 4C m^\varepsilon .
\]
On the \(j\)-th shell,
\[
  \frac{4m^2}{(m^2+s^2)^2}
  \le
  \frac{4m^2}{(2^{2j}m^2)^2},
\]
so \eqref{eq:instanton-u1a-superlinear-mode-count-bound} gives a summable
bound proportional to
\[
  C m^\varepsilon 2^{-(2-\varepsilon)j}
\]
when \(0<\varepsilon<2\); the cases \(\varepsilon\ge2\) are even easier after
enlarging the constant on \(0<\lambda\le\delta\).  The tail above \(\delta\)
is \(O(m^2)\).  Therefore
\begin{equation}
  \Delta_{\pi-\delta}^{\rm zm}(m)\longrightarrow0 .
  \label{eq:instanton-u1a-mode-count-vanishing}
\end{equation}

This is the finite-regulator observable form of the criterion.  A constant
near-zero density has \(N_{\rm zm}(\lambda)\sim\rho_0\lambda\) and produces
the \(1/m\) term in
\eqref{eq:instanton-zero-mode-zone-u1a-kernel}.  A linear density has
\(N_{\rm zm}(\lambda)\sim(c_1/2)\lambda^2\) and can leave a finite remnant.
Vanishing of the Banks--Casher condensate is only the statement that
\(N_{\rm zm}(\lambda)=o(\lambda)\); it is not strong enough.  Suppression of
the \(U(1)_A\)-odd channel requires the stronger cumulative depletion
\eqref{eq:instanton-u1a-superlinear-mode-count-bound}, or an equivalent
bound, after the fluctuation determinant and the ensemble weight have been
included.
\end{controlledapproximation}

For the electroweak \(SU(2)_L\) gauge theory, the same zero-mode logic applies
to every left-handed doublet.  For one generation there are three colored
quark doublets and one lepton doublet, hence four \(SU(2)_L\) doublet zero
modes.  For \(N_g\) generations the corresponding minimal fermion operator
contains
\begin{equation}
  \mathcal O_{\rm EW}
  \propto
  \prod_{r=1}^{N_g}
  \bigl(q_{Lr}q_{Lr}q_{Lr}\ell_{Lr}\bigr),
  \label{eq:electroweak-thooft-vertex}
\end{equation}
with weak, color, spinor, and flavor contractions fixed by the zero-mode
wavefunctions and by gauge invariance.  It changes baryon and lepton numbers
by
\[
  \Delta B=\Delta L=N_g,
  \qquad
  \Delta(B-L)=0,
  \qquad
  \Delta(B+L)=2N_g .
\]
At zero temperature the amplitude is exponentially suppressed by the instanton
action.  The real-time high-temperature process is instead organized around
the sphaleron barrier discussed in
Section~\ref{sec:gauge-higgs-classical-solitons}; both effects are controlled
by the same anomalous relation between topological charge and fermion-number
violation.

\begin{controlledapproximation}[Electroweak \(B+L\) response from topological
transitions]
\label{ca:electroweak-sphaleron-bplusl-response}
The operator in \eqref{eq:electroweak-thooft-vertex} is a zero-temperature
Euclidean tunneling coordinate.  A plasma rate is different data.  In a
thermal \(SU(2)_L\)-Higgs plasma, let
\[
  \Gamma_{\rm CS}
  =
  \lim_{Vt\to\infty}
  \frac{\left\langle
  (N_{\rm CS}(t)-N_{\rm CS}(0))^2\right\rangle}{Vt}
\]
be the Chern--Simons diffusion rate in the normalization where one transition
between neighboring vacua has \(|\Delta N_{\rm CS}|=1\).  The anomaly gives
\begin{equation}
  \Delta B=\Delta L=N_g\,\Delta N_{\rm CS},
  \qquad
  \Delta(B-L)=0,
  \qquad
  \Delta X=2N_g\,\Delta N_{\rm CS},
  \quad X:=B+L .
  \label{eq:electroweak-bplusl-cs-charge-ledger}
\end{equation}
Thus \(B-L\) is not washed out by these transitions, while \(B+L\) is.  If the
long-wavelength plasma free energy for a small homogeneous \(X\)-density is
\[
  f_X(n_X)=\frac{n_X^2}{2\chi_X},
  \qquad
  \mu_X=\frac{\partial f_X}{\partial n_X}=\frac{n_X}{\chi_X},
\]
then detailed balance for diffusion in \(N_{\rm CS}\) gives the linear response
\begin{equation}
  \frac{\dd n_X}{\dd t}
  =
  -
  \frac{(2N_g)^2\Gamma_{\rm CS}}{2\chi_X T}\,n_X .
  \label{eq:electroweak-sphaleron-bplusl-relaxation}
\end{equation}
Again the factor \(1/2\) follows from defining \(\Gamma_{\rm CS}\) as the
variance rate of \(N_{\rm CS}\), not as the Fokker--Planck diffusion
coefficient itself.
The coefficient \(\chi_X\) is the susceptibility of the actual light plasma
degrees of freedom after any imposed conserved charges have been projected out.
It is not determined by the instanton zero-mode counting.  Likewise
\(\Gamma_{\rm CS}\) is a real-time transport coefficient.  In the symmetric
high-temperature phase it is a magnetic-sector rate; in the broken phase it is
Boltzmann-suppressed by the sphaleron barrier.  In a leading-barrier
coordinate one writes
\[
  \Gamma_{\rm CS}^{\rm br}
  =
  \kappa_{\rm br}(T)\,\exp[-E_{\rm sph}(T)/T],
\]
after the negative mode, gauge zero modes, fluctuation determinant, and thermal
prefactor have been fixed.  The Euclidean instanton vertex supplies the
selection rule and the zero-temperature tunneling coordinate; the physical
baryon/lepton washout rate requires the separate sphaleron diffusion problem
and the plasma susceptibility.
\end{controlledapproximation}

\begin{controlledapproximation}[Constrained susceptibility entering
electroweak washout]
\label{ca:electroweak-constrained-washout-susceptibility}
The susceptibility in
\eqref{eq:electroweak-sphaleron-bplusl-relaxation} is itself an observable
thermodynamic input.  Let \(n_i\) be a finite list of slow charge densities in
the plasma, with quadratic free-energy density
\begin{equation}
  f(n)=\frac12 n_i(\chi^{-1})_{ij}n_j,
  \label{eq:electroweak-charge-free-energy-matrix}
\end{equation}
where \(\chi\) is the static susceptibility matrix after fast reactions have
been equilibrated.  Let \(X=w_i n_i\) be the charge coordinate washed out by
sphalerons, and impose independent exactly conserved charges or neutrality
conditions \(C_{\alpha i}n_i=0\), with \(C\chi C^T\) invertible on the
constraint space.  The susceptibility of \(X\) on the constrained thermal
subspace is the Schur complement
\begin{equation}
  \chi_{X\mid C}
  =
  w^T\chi w
  -
  w^T\chi C^T
  (C\chi C^T)^{-1}
  C\chi w .
  \label{eq:electroweak-constrained-susceptibility}
\end{equation}
Equivalently, \(\chi_{X\mid C}\) is defined by minimizing
\eqref{eq:electroweak-charge-free-energy-matrix} at fixed \(X\) and fixed
conserved charges:
\[
  f_{\rm min}(X)=\frac{X^2}{2\chi_{X\mid C}}.
\]
Therefore
\eqref{eq:electroweak-sphaleron-bplusl-relaxation} becomes
\[
  \frac{\dd n_X}{\dd t}
  =
  -
  \frac{(2N_g)^2\Gamma_{\rm CS}}{2\chi_{X\mid C}T}\,n_X
\]
when the only remaining slow departure from equilibrium is this constrained
\(B+L\) coordinate.

For two charges \(n=(n_B,n_L)\), diagonal susceptibilities
\(\chi_B,\chi_L\), and fixed \(B-L\), this gives
\begin{equation}
  \chi_{B+L\mid B-L}
  =
  \chi_B+\chi_L
  -
  \frac{(\chi_B-\chi_L)^2}{\chi_B+\chi_L}
  =
  \frac{4\chi_B\chi_L}{\chi_B+\chi_L}.
  \label{eq:electroweak-bl-projected-susceptibility}
\end{equation}
The naive \(\chi_B+\chi_L\) is correct only if \(B-L\) is allowed to fluctuate
along with \(B+L\).  Conversely, if the imposed constraints already fix
\(B+L\), the Schur complement vanishes and there is no washout coordinate for
the sphaleron diffusion to relax.  This is another place where the Euclidean
't Hooft vertex is not enough: the anomaly supplies the charge step, but the
rate coefficient is set by real-time diffusion and the constrained equilibrium
thermodynamics of the plasma.
\end{controlledapproximation}

\section{Anomalous Pion Couplings}

The axial anomaly has observable low-energy consequences.  In two-flavor QCD,
the neutral pion couples to the electromagnetic topological density through
the anomalous divergence of the axial isospin current:
\[
  \partial_\mu J_{5,3}^\mu
  =
  \frac{e^2}{16\pi^2}
  \operatorname{tr}\bigl(T_3 Q^2\bigr)
  \epsilon^{\mu\nu\rho\sigma}
  F_{\mu\nu}F_{\rho\sigma}
  +
  \text{mass terms}.
\]
Here \(Q\) is the quark charge matrix and \(T_3\) is the third generator of
flavor \(SU(2)\).  The same anomaly is reproduced in the pion effective theory
by the Wess--Zumino--Witten term introduced in the next chapter.
