\chapter{Algebraic QFT: Local Nets and States}
\label{ch:volume-iv-aqft-local-nets-states}
\ConventionsLorentzian


\section{Local Observable Nets}

Let \(M=\mathbb M^D\) be Minkowski spacetime, and let
\(\mathcal K(M)\) be a chosen class of open spacetime regions, for example
double cones or bounded causally complete regions.  The connected Poincare
group is denoted
\[
  \Poinc=\mathbb R^D\rtimes SO^+(1,D-1).
\]
If \(\mathcal O_1,\mathcal O_2\in\mathcal K(M)\), the notation
\(\mathcal O_1\perp\mathcal O_2\) means that every point of
\(\mathcal O_1\) is spacelike separated from every point of
\(\mathcal O_2\).  If \(N\subset M\), \(D(N)\) denotes its domain of
dependence.

\begin{definition}[Haag--Kastler \(C^*\)-net]
\label{def:haag-kastler-cstar-net}
A Haag--Kastler \(C^*\)-net on \(M\), indexed by \(\mathcal K(M)\), consists
of the following data.
\begin{enumerate}[label=\textup{(\roman*)}]
\item For each \(\mathcal O\in\mathcal K(M)\), a unital \(C^*\)-algebra
\(\Obs(\mathcal O)\).
\item For every inclusion \(\mathcal O_1\subset\mathcal O_2\), an injective
unital \(*\)-homomorphism
\(\iota_{\mathcal O_2\mathcal O_1}:\Obs(\mathcal O_1)\to\Obs(\mathcal O_2)\)
compatible with composition.  After using these embeddings, we write
\(\Obs(\mathcal O_1)\subset\Obs(\mathcal O_2)\).
\item A strongly continuous action \(g\mapsto\alpha_g\) of \(\Poinc\) by
\(*\)-automorphisms of the quasilocal algebra
\[
  \Obs_{\rm qloc}
  =
  \overline{\bigcup_{\mathcal O\in\mathcal K(M)}
  \Obs(\mathcal O)}^{\,\|\cdot\|}.
\]
\end{enumerate}
The data obey:
\begin{enumerate}[label=\textup{(HK\arabic*)}]
\item \emph{Isotony:} the inclusions are the embeddings in (ii).
\item \emph{Einstein causality:} if
\(\mathcal O_1\perp\mathcal O_2\), then
\[
  [A,B]=0,\qquad
  A\in\Obs(\mathcal O_1),\quad B\in\Obs(\mathcal O_2).
\]
\item \emph{Poincare covariance:}
\[
  \alpha_g(\Obs(\mathcal O))=\Obs(g\mathcal O),
  \qquad g\in\Poinc .
\]
\item \emph{Time-slice property:} whenever \(N\in\mathcal K(M)\) contains a
Cauchy surface for \(D(N)\) and \(D(N)\in\mathcal K(M)\), the inclusion
\[
  \Obs(N)\longrightarrow\Obs(D(N))
\]
is an isomorphism.
\item \emph{Additivity:} if
\(\mathcal O=\bigcup_i\mathcal O_i\) with
\(\mathcal O_i,\mathcal O\in\mathcal K(M)\), then
\(\Obs(\mathcal O)\) is the \(C^*\)-subalgebra generated by the images of the
\(\Obs(\mathcal O_i)\).
\end{enumerate}
\end{definition}

The definition is the \(C^*\)-algebraic form of the Haag--Kastler locality
framework \cite{HaagKastler1964}.  Its axioms are stated here as data, not as
a consequence of a field presentation; the relation to Wightman fields is a
separate conditional construction below.

Throughout this chapter \(\Obs(\mathcal O)\) denotes the abstract
\(C^*\)-local algebra generated by observables localized inside
\(\mathcal O\), before a Hilbert-space representation has been chosen.  The
norm closure in \(\Obs_{\rm qloc}\) is part of the abstract
\(C^*\)-algebraic presentation.  Weak operator closures belong to represented
von Neumann algebras and are introduced only after a state has selected a
representation.

Figure~\ref{fig:aqft-local-net} records the distinction made in this definition:
the abstract local net assigns \(C^*\)-algebras to regions before any
Hilbert-space state or weak operator closure has been chosen.

\begin{figure}[H]
  \centering
  \resizebox{0.94\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      region/.style={draw,ellipse,line width=0.9pt,fill=blue!5,
        minimum width=2.4cm,minimum height=1.25cm},
      alg/.style={draw,rounded corners=4pt,line width=0.8pt,fill=orange!7,
        minimum width=2.6cm,minimum height=0.75cm,align=center},
      arrow/.style={-{Latex[length=2mm]},line width=0.85pt}
    }
    \begin{scope}[shift={(-3.85,1.25)}]
      \node[region,minimum width=3.35cm,minimum height=1.75cm] (o2) at (0,0) {};
      \node[region,minimum width=1.65cm,minimum height=0.88cm] (o1) at (-0.25,-0.03) {};
      \node at (-0.42,-0.03) {\(\mathcal O_1\)};
      \node at (1.02,0.58) {\(\mathcal O_2\)};
      \node[below] at (0,-1.02) {\(\mathcal O_1\subset\mathcal O_2\)};
    \end{scope}
    \node[alg] (a1) at (0.35,1.52) {\(\Obs(\mathcal O_1)\)};
    \node[alg] (a2) at (3.80,1.52) {\(\Obs(\mathcal O_2)\)};
    \draw[arrow] (-2.10,1.52)--(a1.west) node[midway,above] {assignment};
    \draw[arrow] (a1)--(a2)
      node[midway,above] {\(\iota_{\mathcal O_2\mathcal O_1}\)};
    \node[align=center] at (2.08,0.82)
      {injective \(*\)-homomorphism\\implementing isotony};

    \node[region] (s1) at (-2.55,-1.50) {\(\mathcal O\)};
    \node[region] (s2) at (1.40,-1.50) {\(\widetilde{\mathcal O}\)};
    \draw[densely dashed,line width=0.75pt] (-0.48,-2.15)--(-0.48,-0.85);
    \node[above] at (-0.48,-0.82) {spacelike separated};
    \node[alg] (b1) at (-2.55,-2.95) {\(\Obs(\mathcal O)\)};
    \node[alg] (b2) at (1.40,-2.95) {\(\Obs(\widetilde{\mathcal O})\)};
    \draw[arrow] (s1)--(b1);
    \draw[arrow] (s2)--(b2);
    \draw[<->,line width=0.85pt] (b1)--(b2)
      node[midway,above] {\([A,B]=0\)};
  \end{tikzpicture}
  }
  \caption{The local net is an isotone assignment from suitable spacetime
  regions to abstract \(C^*\)-algebras.  Inclusion of regions is represented
  by the isotony morphism \(\iota_{\mathcal O_2\mathcal O_1}\), while
  spacelike separation is represented algebraically by commutation of the
  corresponding observable algebras.}
  \label{fig:aqft-local-net}
\end{figure}

\paragraph{Benchmark model: the massive scalar Weyl net.}
Before entering perturbative or constructive interacting examples, it is
useful to have one completely explicit local net in view.  Let
\[
  P_m=\partial^\mu\partial_\mu-m^2
\]
be the mostly-plus Klein--Gordon operator on \(M=\mathbb M^D\), with
\(m>0\), and let
\[
  \Delta_m^{\rm ret},\qquad \Delta_m^{\rm adv}
\]
be the retarded and advanced fundamental solutions:
\[
  P_m\Delta_m^{\rm ret/adv}f=f,\qquad
  \Delta_m^{\rm ret/adv}P_m f=f,\qquad
  \operatorname{supp}\Delta_m^{\rm ret/adv}f
  \subset J^\pm(\operatorname{supp}f)
\]
for every \(f\in C_c^\infty(M;\mathbb R)\).  The causal propagator is
\[
  \Delta_m=\Delta_m^{\rm ret}-\Delta_m^{\rm adv}.
\]
Green hyperbolicity gives the exact sequence
\[
  C_c^\infty(M;\mathbb R)
  \xrightarrow{\ P_m\ }
  C_c^\infty(M;\mathbb R)
  \xrightarrow{\ \Delta_m\ }
  C^\infty_{\rm sc}(M;\mathbb R),
\]
where \(C^\infty_{\rm sc}\) denotes smooth solutions with spatially compact
support.  Exactness at the middle term says
\[
  \ker\Delta_m=P_m C_c^\infty(M;\mathbb R).
\]
Thus the real vector space
\begin{equation}
  \mathsf V_m
  :=
  C_c^\infty(M;\mathbb R)\big/P_m C_c^\infty(M;\mathbb R)
  \label{eq:free-scalar-symplectic-test-quotient}
\end{equation}
is the test-function coordinatization of the classical solution space.  For
\([f],[g]\in\mathsf V_m\) define
\begin{equation}
  \sigma_m([f],[g])
  =
  \int_M f(x)\,(\Delta_m g)(x)\,\dd^D x .
  \label{eq:free-scalar-symplectic-form}
\end{equation}
This is well defined because \(P_m\) is formally self-adjoint and
\(P_m\Delta_m=\Delta_m P_m=0\) on test functions.  It is antisymmetric since
\(\Delta_m(x,y)=-\Delta_m(y,x)\).  It is nondegenerate on the quotient:
if \(\sigma_m([f],[g])=0\) for all \(g\), then the distribution
\(\Delta_m f\) pairs trivially with all compactly supported test functions,
so \(\Delta_m f=0\), hence \(f\in P_m C_c^\infty(M;\mathbb R)\) by exactness.

The Weyl \(C^*\)-algebra
\[
  \mathfrak W(\mathsf V_m,\sigma_m)
\]
is the universal \(C^*\)-algebra generated by symbols \(W(u)\),
\(u\in\mathsf V_m\), subject to
\begin{equation}
  W(0)=1,\qquad W(u)^*=W(-u),\qquad
  W(u)W(v)=\ee^{-\frac{\ii}{2}\sigma_m(u,v)}W(u+v).
  \label{eq:free-scalar-weyl-relations}
\end{equation}
The sign in \eqref{eq:free-scalar-weyl-relations} matches the field
normalization of Volume~I: if \(W([f])=\exp(\ii\widehat\phi(f))\) in the Fock
representation, then
\[
  [\widehat\phi(f),\widehat\phi(g)]
  =
  \ii\sigma_m([f],[g])\,1,
\]
and the Baker--Campbell--Hausdorff formula gives the displayed Weyl relation.
For an open region \(\mathcal O\), let
\[
  \mathsf V_m(\mathcal O)
  =
  \operatorname{image}\bigl(C_c^\infty(\mathcal O;\mathbb R)
  \longrightarrow\mathsf V_m\bigr),
\]
and define
\begin{equation}
  \Obs_m(\mathcal O)
  =
  C^*\bigl(W(u)\mid u\in\mathsf V_m(\mathcal O)\bigr)
  \subset\mathfrak W(\mathsf V_m,\sigma_m).
  \label{eq:free-scalar-weyl-local-algebra}
\end{equation}

The Haag--Kastler properties can now be checked without appealing to
unbounded fields.  Isotony follows immediately from
\(C_c^\infty(\mathcal O_1)\subset C_c^\infty(\mathcal O_2)\).  If
\(\mathcal O_1\perp\mathcal O_2\), then
\(\operatorname{supp}\Delta_m g\subset J(\operatorname{supp}g)\) is disjoint
from \(\operatorname{supp}f\) for
\(f\in C_c^\infty(\mathcal O_1)\),
\(g\in C_c^\infty(\mathcal O_2)\).  Hence
\(\sigma_m([f],[g])=0\), and the Weyl relations give
\[
  W([f])W([g])=W([g])W([f]).
\]
Since the local algebras are \(C^*\)-closures of the algebras generated by
such Weyl elements, spacelike separated local algebras commute.  Poincare
covariance is induced by
\[
  \alpha_g\bigl(W([f])\bigr)=W([f\circ g^{-1}]),
  \qquad g\in\Poinc,
\]
where \(g\) acts on test functions by pullback.  The covariance of
\(P_m\), \(\Delta_m^{\rm ret/adv}\), and the volume form preserves the
quotient and the symplectic form.

Additivity follows from compact support.  Suppose
\(\mathcal O=\bigcup_i\mathcal O_i\), and let
\(f\in C_c^\infty(\mathcal O;\mathbb R)\).  The compact set
\(\operatorname{supp}f\) is covered by finitely many \(\mathcal O_i\)'s.
Choose a smooth partition of unity \(\{\rho_j\}_{j=1}^r\) subordinate to this
finite cover on a neighborhood of \(\operatorname{supp}f\).  Then
\[
  f=\sum_{j=1}^r f_j,\qquad f_j=\rho_j f,\qquad
  \operatorname{supp}f_j\subset\mathcal O_{i_j}.
\]
In \(\mathsf V_m(\mathcal O)\) this gives
\([f]=\sum_j[f_j]\).  Iterating the Weyl relation
\eqref{eq:free-scalar-weyl-relations} expresses \(W([f])\), up to an
explicit phase determined by the pairings
\(\sigma_m([f_j],[f_k])\), as a product of the \(W([f_j])\)'s.  Hence every
generator of \(\Obs_m(\mathcal O)\) belongs to the \(C^*\)-algebra generated
by the \(\Obs_m(\mathcal O_i)\)'s, and the opposite inclusion is isotony.
The phase just mentioned is not an additional analytic hypothesis.  In any
finite-dimensional symplectic vector space \((V_{\rm fin},\sigma)\) with Weyl
relations
\[
  W(u)W(v)=\exp\!\left(-\frac{\ii}{2}\sigma(u,v)\right)W(u+v),
\]
one has
\[
  W(u)W(v)=\exp[-\ii\sigma(u,v)]\,W(v)W(u),
  \qquad
  W(u)W(v)W(u)^*W(v)^*
  =
  \exp[-\ii\sigma(u,v)]\,1.
\]
Thus vanishing of the symplectic pairing is exactly the finite algebraic
mechanism behind local commutation.  Likewise, for a finite decomposition
\(u=\sum_{j=1}^r u_j\),
\[
  W(u)
  =
  \exp\!\left(\frac{\ii}{2}\sum_{1\le j<k\le r}\sigma(u_j,u_k)\right)
  W(u_1)\cdots W(u_r).
\]
The continuum additivity proof above supplies the support-theoretic
decomposition \( [f]=\sum_j[f_j]\) in the Green-hyperbolic quotient; the
displayed finite cocycle identity is the bounded-algebra step that converts
that decomposition into containment of local \(C^*\)-algebras.

The time-slice property is the only axiom in this model that uses a real
hyperbolic argument.  Let \(U=D(N)\), and assume the chosen region class is
regular enough that \(N\subset U\) contains a Cauchy-surface collar: there is
a smooth Cauchy time function \(\tau\) on \(U\) and numbers
\(\tau_-<\tau_+\) for which
\(\tau^{-1}([\tau_-,\tau_+])\subset N\).  Choose
\(\chi\in C^\infty(U)\) with \(\chi=0\) for \(\tau<\tau_-\) and
\(\chi=1\) for \(\tau>\tau_+\).  For \(f\in C_c^\infty(U;\mathbb R)\), set
\[
  \varphi=\Delta_m f,\qquad
  f_N=P_m(\chi\varphi).
\]
Because \(P_m\varphi=0\), the support of \(f_N\) lies where derivatives of
\(\chi\) are nonzero, hence in \(N\).  To compare the quotient classes, note
that \(\chi\varphi\) vanishes in the far past and solves
\[
  P_m(\chi\varphi)=f_N;
\]
by uniqueness of the retarded solution it is \(\Delta_m^{\rm ret}f_N\).
Likewise \((\chi-1)\varphi\) vanishes in the far future and solves
\[
  P_m\bigl((\chi-1)\varphi\bigr)=f_N;
\]
by uniqueness of the advanced solution it is \(\Delta_m^{\rm adv}f_N\).
Therefore
\[
  \Delta_m f_N
  =
  \Delta_m^{\rm ret}f_N-\Delta_m^{\rm adv}f_N
  =
  \chi\varphi-(\chi-1)\varphi
  =
  \varphi
  =
  \Delta_m f .
\]
Exactness of the Green-hyperbolic sequence then gives
\([f]=[f_N]\) in the quotient associated with \(U\).  Since
\(\mathsf V_m(N)\) and \(\mathsf V_m(U)\) were defined as image subspaces of
the global quotient, this proves equality of image subspaces,
\[
  \mathsf V_m(N)=\mathsf V_m(U).
\]
The generated Weyl algebras are
therefore identical:
\[
  \Obs_m(N)=\Obs_m(D(N)).
\]

The usual vacuum representation is a state on this abstract net, not part of
the net definition.  The positive-frequency two-point distribution
\(\Delta_+\) defines a real covariance
\[
  \mu_m([f],[g])=\operatorname{Re}
  \int_{M\times M}f(x)\Delta_+(x-y)g(y)\,\dd^D x\,\dd^D y
\]
satisfying
\[
  |\sigma_m(u,v)|^2\le 4\,\mu_m(u,u)\mu_m(v,v).
\]
Consequently
\[
  \omega_0(W(u))=\exp\!\left[-\frac12\mu_m(u,u)\right]
\]
is a regular quasi-free state, and its GNS representation is the scalar Fock
representation.  This final step illustrates the structural separation:
\(\mathcal O\mapsto\Obs_m(\mathcal O)\) is the abstract observable net, while
\((\Hilb_{\omega_0},\pi_{\omega_0},\Omega_{\omega_0})\) is selected only
after a state is chosen.

\section{Perturbative Algebraic QFT and Interacting Examples}
\label{sec:perturbative-aqft-bridge}

Perturbative algebraic QFT is the framework in which Epstein--Glaser causal
renormalization, local time-ordered products, and algebraic locality are put
in the same language.  Its algebras are formal power series in the
interaction and in \(\hbar\); at every fixed formal order the locality and
renormalization statements are statements about distributions and their
extensions.

\begin{definition}[Microcausal functionals and the BDF star product]
\label{def:perturbative-algebraic-qft-star-product}
For a scalar field on a globally hyperbolic spacetime \(M\), let
\(\mathcal E=C^\infty(M)\) be the classical configuration space.  A smooth
functional \(F:\mathcal E\to\mathbb C\) belongs to
\(\mathcal F_{\mu{\rm c}}\), the space of microcausal functionals, if every
functional derivative \(F^{(n)}(\varphi)\) is a compactly supported
distribution on \(M^n\) and
\begin{equation}
  \operatorname{WF}\bigl(F^{(n)}(\varphi)\bigr)
  \cap
  \bigl(\overline V_+^n\cup\overline V_-^n\bigr)
  =
  \emptyset .
  \label{eq:microcausal-wavefront-condition}
\end{equation}
Here \(\overline V_\pm^n\) is the set of \(n\)-tuples of nonzero covectors all
lying in the closed future or all lying in the closed past light cone.  If
\(H(x,y)\) is a Hadamard bidistribution for the free theory, the
Brunetti--Duetsch--Fredenhagen product is
\begin{equation}
  F\star_H G
  =
  m\circ
  \exp\!\left[
    \hbar
    \int_{M^2}
      H(x,y)
      \frac{\delta}{\delta\varphi(x)}
      \otimes
      \frac{\delta}{\delta\varphi(y)}
      \,\dd x\,\dd y
  \right](F\otimes G),
  \label{eq:paqft-star-product}
\end{equation}
where \(m(F\otimes G)(\varphi)=F(\varphi)G(\varphi)\).  The exponential is
expanded as a formal power series and every term is defined by the
Hörmander wavefront-set product criterion applied to
\eqref{eq:microcausal-wavefront-condition} and the Hadamard wavefront set of
\(H\).
\end{definition}

\paragraph{Peierls bracket from the pAQFT star commutator.}
Assume that the antisymmetric part of \(H\) is \((\ii/2)\Delta\), where
\(\Delta\) is the causal propagator of the free equation of motion.  For
regular functionals, and then by the microcausal extension wherever the
products are defined,
\begin{equation}
  [F,G]_{\star_H}
  =
  \ii\hbar\,\{F,G\}_{\rm Peierls}
  +O(\hbar^2).
  \label{eq:paqft-star-commutator-peierls}
\end{equation}
The order-\(\hbar\) part of
\eqref{eq:paqft-star-product} is
\[
  \hbar\int_{M^2} H(x,y)
  \frac{\delta F}{\delta\varphi(x)}
  \frac{\delta G}{\delta\varphi(y)}\,\dd x\,\dd y .
\]
Subtracting the same expression with \(F\) and \(G\) interchanged replaces
\(H(x,y)\) by \(H(x,y)-H(y,x)=\ii\Delta(x,y)\).  The resulting bilinear form
is exactly \(\ii\hbar\{F,G\}_{\rm Peierls}\) in the free theory.  Higher terms
in the exponential contribute \(O(\hbar^2)\).

\begin{proposition}[Perturbative interacting local net]
\label{prop:relative-s-matrices-give-perturbative-local-net}
Let \(T_n:\mathcal F_{\rm loc}^{\otimes n}\to\mathcal F_{\mu{\rm c}}\) be
Epstein--Glaser time-ordered products satisfying symmetry, covariance,
microlocal spectrum bounds, unitarity, field independence, and causal
factorization.  Define
\[
  \mathcal S(F)
  =
  1+\sum_{n\ge1}\frac{\ii^n}{\hbar^n n!}\,T_n(F^{\otimes n}).
\]
Causal factorization is the identity
\begin{equation}
  \mathcal S(F+G)=\mathcal S(F)\star_H\mathcal S(G)
  \quad\text{if}\quad
  \operatorname{supp}F\cap J^-(\operatorname{supp}G)=\emptyset .
  \label{eq:paqft-causal-factorization}
\end{equation}
We shall also use its three-functional form
\begin{equation}
  \mathcal S(F+G+H)
  =
  \mathcal S(F+G)\star_H
  \mathcal S(G)^{-1}_{\star_H}\star_H
  \mathcal S(G+H),
  \qquad
  \operatorname{supp}F\cap J^-(\operatorname{supp}H)=\emptyset ,
  \label{eq:paqft-causal-factorization-three-term}
\end{equation}
which is part of the causal-factorization axiom for the formal \(S\)-matrix.
For an interaction functional \(V\), set
\begin{equation}
  \mathcal S_V(F)
  :=
  \mathcal S(V)^{-1}_{\star_H}\star_H \mathcal S(V+F),
  \label{eq:paqft-relative-s-matrix}
\end{equation}
where the inverse is taken in the completed formal \(\star_H\)-algebra.  Let
\(\mathfrak A_V^{\rm pert}(\mathcal O)\) be the formal \(*\)-algebra
generated by \(\mathcal S_V(F)\) with \(\operatorname{supp}F\subset\mathcal O\),
modulo the causal-factorization relations and the time-slice ideal.  Then
\[
  \mathcal O\longmapsto \mathfrak A_V^{\rm pert}(\mathcal O)
\]
is an isotonic, time-slice local net of formal algebras.  Covariance has the
following precise form: a spacetime symmetry \(g\) sends
\(\mathfrak A_V^{\rm pert}(\mathcal O)\) to
\(\mathfrak A_{gV}^{\rm pert}(g\mathcal O)\); if \(V\) is invariant under the
symmetry, this is covariance of a single net.  Locality means that generators
with spacelike separated supports commute.
\end{proposition}

\begin{proof}
The formal inverse in \eqref{eq:paqft-relative-s-matrix} exists because
\(\mathcal S(V)=1+O(V)\) in the completed formal algebra.  The net maps for
inclusions are the homomorphisms induced by the identity on generators: if
\(\mathcal O_1\subset\mathcal O_2\), every generator with support in
\(\mathcal O_1\) is also one with support in \(\mathcal O_2\), and the
relations imposed in the smaller algebra are restrictions of the relations
imposed in the larger algebra.  This proves isotony.  The time-slice property
is exactly the quotient by the time-slice ideal: if
\(\mathcal N\subset\mathcal O\) contains a Cauchy surface for
\(\mathcal O\), the ideal identifies the generators in \(\mathcal O\) with
the algebra generated by representatives supported in \(\mathcal N\), using
the free equation of motion and causal propagation in the Epstein--Glaser
construction.

For covariance, let \(g\) be a spacetime symmetry preserving the background
data used in \(H\).  Covariance of the time-ordered products gives
\[
  \alpha_g(\mathcal S(F))=\mathcal S(gF),
  \qquad
  \alpha_g(A\star_H B)=\alpha_g(A)\star_H\alpha_g(B),
\]
where \(gF\) is the pushforward of the local functional.  Applying this to
\eqref{eq:paqft-relative-s-matrix} gives
\[
  \alpha_g\bigl(\mathcal S_V(F)\bigr)
  =
  \mathcal S_{gV}(gF),
  \qquad
  \operatorname{supp}(gF)=g(\operatorname{supp}F).
\]
Thus \(g\) sends the algebra of \(\mathcal O\) for \(V\) to the algebra of
\(g\mathcal O\) for the transformed interaction \(gV\).  If \(gV=V\), this is
a covariance automorphism of the same net.

It remains to prove locality.  We first record the relative causal
factorization identity.  If
\[
  \operatorname{supp}F\cap J^-(\operatorname{supp}G)=\emptyset ,
\]
then \eqref{eq:paqft-causal-factorization-three-term}, applied with the
interaction \(V\) as the middle functional, gives
\[
  \mathcal S(V+F+G)
  =
  \mathcal S(V+F)\star_H
  \mathcal S(V)^{-1}_{\star_H}\star_H
  \mathcal S(V+G).
\]
Multiplying on the left by \(\mathcal S(V)^{-1}_{\star_H}\) yields
\begin{equation}
  \mathcal S_V(F+G)
  =
  \mathcal S_V(F)\star_H\mathcal S_V(G).
  \label{eq:relative-s-causal-factorization-proof}
\end{equation}
For spacelike separated supports both causal orderings are allowed:
\[
  \operatorname{supp}F\cap J^-(\operatorname{supp}G)=\emptyset,
  \qquad
  \operatorname{supp}G\cap J^-(\operatorname{supp}F)=\emptyset .
\]
Using \eqref{eq:relative-s-causal-factorization-proof} in the two orders and
using symmetry of the time-ordered products,
\(\mathcal S_V(F+G)=\mathcal S_V(G+F)\), gives
\[
  \mathcal S_V(F)\star_H\mathcal S_V(G)
  =
  \mathcal S_V(G)\star_H\mathcal S_V(F).
\]
Since the generators of
\(\mathfrak A_V^{\rm pert}(\mathcal O_1)\) and
\(\mathfrak A_V^{\rm pert}(\mathcal O_2)\) are of this form, and the quotient
relations are compatible with \(\star_H\), this proves the locality relation
for spacelike separated regions.
\end{proof}

Renormalization in this framework is the freedom in extending \(T_n\) to the
diagonals.  The ambiguity is local: two choices are related by a
Stueckelberg--Petermann redefinition of the interaction by local functionals
and their derivatives.  For gauge theory the same construction must be
combined with the BV complex and the quantum master equation; the physical
perturbative algebra is then the BRST/BV cohomology of the interacting
algebra.  These facts make perturbative AQFT a natural mathematical home for
local counterterms, contact terms, and gauge-theoretic Ward identities.  It
will not be treated only as a dictionary between path integrals and local
nets.  Later chapters must develop the perturbative algebra itself: the
classification of time-ordered-product extensions, the renormalization group
as the Stueckelberg--Petermann group, interacting fields by Bogoliubov's
formula, the time-slice axiom in perturbation theory, BV cohomology for gauge
theories, and the relation between these constructions and the 1PI and
Wilsonian effective actions.

The algebraic formulation must also be developed through substantial
interacting examples.  A persistent weakness of much abstract AQFT literature
is that the axioms, sector theory, and local-algebra technology are developed
with too few nontrivial models in view.  In this monograph, whenever an
interacting construction is available or partially available, the associated
algebraic questions should be asked explicitly: what are the local algebras or
formal local algebras, what is the state space, which sectors are present, how
does scattering or infraparticle structure appear, how are scaling limits and
operator product structures represented, and which steps are theorem-level
rather than expected structure?  Constructive scalar models, perturbatively
constructed Epstein--Glaser/pAQFT models, factorizing two-dimensional models,
lattice-continuum scaling limits, conformal nets, and gauge theories with
BRST/BV control are therefore not examples after the axioms; they are part of
the theory development and tests of whether the algebraic formulation captures
the physics it is meant to organize.

The purpose of including this material is not only to collect known
constructions.  The comparison among Wightman, Euclidean, Wilsonian,
perturbative algebraic, BV, and operator-algebraic formulations exposes
genuine gaps: gauge-theory local nets beyond perturbation theory,
charged sectors requiring nonlocal dressing, infraparticle scattering,
contact-term control in deformations, and the passage from constructive or
lattice scaling limits to detailed algebraic data.  These gaps should be
formulated as precise mathematical and physical problems and, where feasible,
developed within the monograph itself.  Existing results supply boundary
conditions for this development; they are not the endpoint of the project.
No axiomatic framework is taken here as sacred or final.  A framework earns
its place by clarifying nontrivial physics and the mathematics behind it:
which observables are defined, which limits exist, which locality and
positivity statements are actually proven, how sectors and scattering are
encoded, and where the formulation fails to see structures that are physically
present.  The comparison of frameworks is therefore part of the subject
matter, not a preliminary choice of doctrine.

\section{States and GNS Representations}

\begin{definition}[State on a \(C^*\)-algebra]
\label{def:cstar-state}
A state on a unital \(C^*\)-algebra \(\Obs\) is a linear functional
\[
  \omega:\Obs\to\mathbb C
\]
such that
\[
  \omega(A^*A)\ge0,\qquad \omega(1)=1.
\]
\end{definition}

\noindent
The source trace for the operator-algebra form of the construction used here
is the Gelfand--Naimark representation theorem together with Segal's
state-to-representation construction
\cite{GelfandNaimark1943,Segal1947}; Takesaki's account is the modular-theory
reference used later in this volume \cite{Takesaki1970}.

\begin{theorem}[Gelfand--Naimark--Segal construction]
\label{thm:gns-construction}
Let \(\omega\) be a state on a unital \(C^*\)-algebra \(\Obs\).  There are a
Hilbert space \(\Hilb_\omega\), a unital \(*\)-representation
\(\pi_\omega:\Obs\to B(\Hilb_\omega)\), and a cyclic unit vector
\(\Omega_\omega\in\Hilb_\omega\) such that
\begin{equation}
  \omega(A)=
  \langle\Omega_\omega,\pi_\omega(A)\Omega_\omega\rangle_\omega ,
  \qquad A\in\Obs .
  \label{eq:gns-state-vector-functional}
\end{equation}
If \((\Hilb',\pi',\Omega')\) is another cyclic representation satisfying
\eqref{eq:gns-state-vector-functional}, there is a unique unitary
\(W:\Hilb_\omega\to\Hilb'\) with
\[
  W\Omega_\omega=\Omega',
  \qquad
  W\pi_\omega(A)=\pi'(A)W .
\]
\end{theorem}

\begin{proof}
Let
\[
  \mathcal N_\omega=\{A\in\Obs\mid \omega(A^*A)=0\}.
\]
The positivity of \(\omega\) gives the Cauchy--Schwarz inequality
\[
  |\omega(C^*D)|^2\le \omega(C^*C)\,\omega(D^*D),
  \qquad C,D\in\Obs.
\]
Hence \(C\in\mathcal N_\omega\) implies
\(\omega(C^*D)=\omega(D^*C)=0\) for all \(D\in\Obs\), so
\(\mathcal N_\omega\) is a linear subspace.  It is also a left ideal.  If
\(C\in\mathcal N_\omega\) and \(A\in\Obs\), then the \(C^*\)-algebra order
inequality \(0\le A^*A\le \|A\|^2 1\) gives
\[
  0\le \omega((AC)^*(AC))
  =
  \omega(C^*A^*AC)
  \le
  \|A\|^2\omega(C^*C)
  =
  0 .
\]
Thus \(AC\in\mathcal N_\omega\).  The quotient
\(\Obs/\mathcal N_\omega\) therefore has the well-defined inner product
\[
  \langle [A],[B]\rangle_\omega=\omega(A^*B).
\]
Let \(\Hilb_\omega\) be its completion.  Define
\[
  \pi_\omega(A)[B]=[AB],
  \qquad
  \Omega_\omega=[1].
\]
The representation is well-defined because \(\mathcal N_\omega\) is a left
ideal, and
\[
  \|\pi_\omega(A)[B]\|_\omega^2
  =
  \omega(B^*A^*AB)
  \le
  \|A\|^2\omega(B^*B),
\]
so \(\pi_\omega(A)\) extends to a bounded operator on \(\Hilb_\omega\).
Equation \eqref{eq:gns-state-vector-functional} follows from the definition
of the inner product, and \(\Omega_\omega\) is cyclic because
\(\pi_\omega(\Obs)\Omega_\omega\) contains all quotient classes \([A]\).

For uniqueness, define \(W\) on the dense subspace
\(\pi_\omega(\Obs)\Omega_\omega\) by
\[
  W\pi_\omega(A)\Omega_\omega=\pi'(A)\Omega'.
\]
This is well-defined and isometric because both cyclic vector states equal
\(\omega\):
\[
  \langle\pi_\omega(A)\Omega_\omega,\pi_\omega(B)\Omega_\omega\rangle
  =
  \omega(A^*B)
  =
  \langle\pi'(A)\Omega',\pi'(B)\Omega'\rangle .
\]
Cyclicity of \(\Omega'\) makes the range dense, hence \(W\) extends to a
unitary.  The intertwining identity and \(W\Omega_\omega=\Omega'\) follow on
the dense cyclic subspace.
\end{proof}

Figure~\ref{fig:gns-representation} summarizes the GNS construction: a state
defines the null ideal, the quotient pre-Hilbert space, the cyclic vector, and
the represented algebra, uniquely up to unitary equivalence.

\begin{figure}[H]
  \centering
  \resizebox{0.96\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      box/.style={draw,rounded corners=4pt,line width=0.8pt,fill=blue!4,
        minimum width=2.95cm,minimum height=0.88cm,align=center},
      arrow/.style={-{Latex[length=2mm]},line width=0.85pt}
    }
    \node[box] (alg) at (-5.7,1.25) {abstract algebra\\\(\Obs\)};
    \node[box,fill=yellow!10] (state) at (-5.7,-0.55)
      {state functional\\\(\omega:\Obs\to\mathbb C\)};
    \node[box] (null) at (-2.45,1.25)
      {null left ideal\\\(\mathcal N_\omega=\{A:\omega(A^*A)=0\}\)};
    \node[box] (quot) at (1.10,1.25)
      {pre-Hilbert quotient\\\(\Obs/\mathcal N_\omega\)\\
      \(\langle[A],[B]\rangle=\omega(A^*B)\)};
    \node[box] (hilb) at (4.82,1.25)
      {completion\\\(\Hilb_\omega\), \(\Omega_\omega=[1]\)};
    \node[box,fill=green!7] (rep) at (1.10,-1.25)
      {left multiplication\\\(\pi_\omega(A)[B]=[AB]\)};
    \node[box,fill=green!7] (local) at (4.82,-1.25)
      {represented local algebras\\\(\pi_\omega(\Obs(\mathcal O))''\)};
    \draw[arrow] (alg)--(null);
    \draw[arrow] (state)--(null)
      node[midway,below,sloped] {positivity};
    \draw[arrow] (null)--(quot);
    \draw[arrow] (quot)--(hilb);
    \draw[arrow] (alg.south east) to[out=-25,in=180] (rep.west);
    \draw[arrow] (quot)--(rep);
    \draw[arrow] (rep)--(local);
    \draw[arrow] (hilb)--(local);
  \end{tikzpicture}
  }
  \caption{The GNS representation is obtained by quotienting the abstract
  algebra by the null ideal determined by the state, completing the resulting
  inner-product space, and representing the algebra by left multiplication.
  Local \(C^*\)-algebras become represented von Neumann algebras only after
  this state-dependent Hilbert space has been constructed.}
  \label{fig:gns-representation}
\end{figure}

\section{Vacuum Representations}

\begin{definition}[Vacuum state and represented vacuum net]
\label{def:vacuum-representation-aqft}
Let \(\mathcal O\mapsto\Obs(\mathcal O)\) be a Poincare-covariant
Haag--Kastler \(C^*\)-net.  A state \(\omega_0\) on
\(\Obs_{\rm qloc}\) is a vacuum state when:
\begin{enumerate}[label=\textup{(\roman*)}]
\item \(\omega_0\) is Poincare invariant:
\[
  \omega_0(\alpha_g(A))=\omega_0(A),
  \qquad A\in\Obs_{\rm qloc},\quad g\in\Poinc ;
\]
\item in the GNS representation
\((\Hilb_{\omega_0},\pi_{\omega_0},\Omega_{\omega_0})\), translations are
implemented by a strongly continuous unitary representation whose joint
spectrum lies in \(\overline V_+\);
\item \(\Omega_{\omega_0}\) is the unique invariant vector up to phase.
\end{enumerate}
The represented von Neumann local algebra of a region is
\[
  \mathcal R_{\omega_0}(\mathcal O)
  =
  \pi_{\omega_0}(\Obs(\mathcal O))''.
\]
The double commutant indicates weak operator closure.
\end{definition}

\begin{proposition}[Unitary implementation of an invariant state]
\label{prop:invariant-state-unitary-implementation}
Let \(\omega_0\) be invariant under \(\Poinc\).  In its GNS representation
there is a unique strongly continuous unitary representation
\(U_{\omega_0}\) of \(\Poinc\) satisfying
\begin{equation}
  U_{\omega_0}(g)\pi_{\omega_0}(A)U_{\omega_0}(g)^{-1}
  =
  \pi_{\omega_0}(\alpha_g(A)),
  \qquad
  U_{\omega_0}(g)\Omega_{\omega_0}=\Omega_{\omega_0}.
  \label{eq:vacuum-unitary-implementation}
\end{equation}
\end{proposition}

\begin{proof}
On the algebraic cyclic subspace
\(\mathcal D_0=\pi_{\omega_0}(\Obs_{\rm qloc})\Omega_{\omega_0}\), define
\[
  U_{\omega_0}(g)\pi_{\omega_0}(A)\Omega_{\omega_0}
  =
  \pi_{\omega_0}(\alpha_g(A))\Omega_{\omega_0}.
\]
This is well defined on the GNS quotient.  If
\(\pi_{\omega_0}(A)\Omega_{\omega_0}=0\), then
\(\omega_0(A^*A)=0\); invariance gives
\[
  \omega_0(\alpha_g(A)^*\alpha_g(A))
  =
  \omega_0(\alpha_g(A^*A))
  =
  \omega_0(A^*A)=0,
\]
so \(\pi_{\omega_0}(\alpha_g(A))\Omega_{\omega_0}=0\).  The same
calculation shows preservation of inner products:
\[
  \langle\pi_{\omega_0}(\alpha_g(A))\Omega_{\omega_0},
  \pi_{\omega_0}(\alpha_g(B))\Omega_{\omega_0}\rangle
  =
  \omega_0(\alpha_g(A^*B))
  =
  \omega_0(A^*B).
\]
Thus \(U_{\omega_0}(g)\) extends from \(\mathcal D_0\) to an isometry on
\(\Hilb_{\omega_0}\).  Applying the same construction to \(g^{-1}\) gives
the inverse isometry, so \(U_{\omega_0}(g)\) is unitary.  The group law holds
on \(\mathcal D_0\):
\[
  U_{\omega_0}(g)U_{\omega_0}(h)\pi_{\omega_0}(A)\Omega_{\omega_0}
  =
  \pi_{\omega_0}(\alpha_{gh}(A))\Omega_{\omega_0},
\]
and hence holds on the Hilbert-space completion.  The covariance relation in
\eqref{eq:vacuum-unitary-implementation} is checked on
\(\mathcal D_0\) and extended by boundedness.

For strong continuity, it is enough to test vectors in the dense
\(\mathcal D_0\), uniformly boundedness then extends the result to all
vectors.  For \(A\in\Obs_{\rm qloc}\),
\[
  \|U_{\omega_0}(g)\pi_{\omega_0}(A)\Omega_{\omega_0}
      -\pi_{\omega_0}(A)\Omega_{\omega_0}\|^2
  =
  \omega_0\!\left(
    (\alpha_g(A)-A)^*(\alpha_g(A)-A)
  \right),
\]
which tends to zero by the assumed strong continuity of the automorphic
action in the GNS norm.  Finally, if a unitary representation satisfying
\eqref{eq:vacuum-unitary-implementation} and fixing \(\Omega_{\omega_0}\)
exists, its value on every vector
\(\pi_{\omega_0}(A)\Omega_{\omega_0}\) is forced by the covariance relation;
density gives uniqueness.
\end{proof}

When the vacuum representation is fixed, we write \(\mathcal R(\mathcal O)\)
for \(\mathcal R_{\omega_0}(\mathcal O)\).  Thus, in the remainder of this
chapter, \(\Obs(\mathcal O)\) denotes the abstract local \(C^*\)-algebra,
whereas \(\mathcal R(\mathcal O)\) denotes its vacuum-representation von
Neumann completion \(\pi_{\omega_0}(\Obs(\mathcal O))''\).  These are related
by the chosen representation and are distinct mathematical objects.

\section{Correlation Functions in a Representation}

The algebraic presentation produces correlation functions once a state and a
representation have been chosen.  If
\[
  A_j\in\Obs(\mathcal O_j),
  \qquad j=1,\ldots,n,
\]
are bounded local observables, their vacuum correlation function is
\[
  \omega_0(A_1\cdots A_n)
  =
  \langle\Omega_{\omega_0},
  \pi_{\omega_0}(A_1)\cdots\pi_{\omega_0}(A_n)
  \Omega_{\omega_0}\rangle .
\]
Covariance gives spacetime dependence by translating the local observables:
\[
  A_j(x_j)=\alpha_{x_j}(A_j).
\]
Thus bounded local observables give well-defined functions or distributions
after smearing in the spacetime parameters \(x_j\), depending on the
continuity and phase-space properties imposed on the net.

Point fields require additional structure.  A field \(\Phi(f)\) localized in
\(\mathcal O\) should be affiliated with the represented algebra
\(\mathcal R(\mathcal O)\), or
should arise as a limit of bounded local observables in a topology specified
by the construction.  When this is available, the Wightman distributions are
recovered as vacuum matrix elements of these affiliated unbounded operators.
The algebraic net therefore supplies the local observable content, while
field coordinatizations specify distributional generators for that content.

\section{Haag Duality and the Dual Net}

\begin{definition}[Causal complement and Haag duality]
\label{def:haag-duality}
For a represented net, the causal complement of a region \(\mathcal O\) is
\[
  \mathcal O^\perp
  =
  \{x\in M:\ (x-y)^2>0\ \text{for all }y\in\mathcal O\}.
\]
The net satisfies Haag duality for \(\mathcal O\) if
\begin{equation}
  \mathcal R(\mathcal O)=\mathcal R(\mathcal O^\perp)' .
  \label{eq:haag-duality}
\end{equation}
It says that every bounded operator commuting with all observables localized
outside \(\mathcal O\) is already generated by observables localized in
\(\mathcal O\).
\end{definition}

Let \(\mathcal R\) be a local represented net indexed by causally complete
regions, so \(\mathcal O=\mathcal O^{\perp\perp}\).  Define
\[
  \mathcal R^d(\mathcal O):=\mathcal R(\mathcal O^\perp)' .
\]
Then \(\mathcal R^d\) is isotonic and
\[
  \mathcal R(\mathcal O)\subset\mathcal R^d(\mathcal O).
\]
The equality \(\mathcal R=\mathcal R^d\) is precisely Haag duality.  The
locality of \(\mathcal R^d\) is an additional property, often called
essential duality.
Einstein causality gives
\[
  \mathcal R(\mathcal O)\subset \mathcal R(\mathcal O^\perp)'
  =
  \mathcal R^d(\mathcal O).
\]
If \(\mathcal O_1\subset\mathcal O_2\), then
\(\mathcal O_2^\perp\subset\mathcal O_1^\perp\), hence
\(\mathcal R(\mathcal O_2^\perp)\subset\mathcal R(\mathcal O_1^\perp)\).
Taking commutants reverses the inclusion and gives
\(\mathcal R^d(\mathcal O_1)\subset\mathcal R^d(\mathcal O_2)\).  This is
isotony.

The inclusion \(\mathcal R\subset\mathcal R^d\) measures the possible gap
between the chosen local observable generators and the maximal algebra
determined by causal commutation.  In superselection theory, Haag duality is
the condition that permits localized representations to be represented
cleanly by localized endomorphisms of the observable algebra.

\paragraph{Dual-net algebra.}
The algebra \(\mathcal R^d(\mathcal O)\) is the algebra detected by all
causal-commutation tests outside \(\mathcal O\).  Thus an operator
\(T\in B(\Hilb)\) belongs to \(\mathcal R^d(\mathcal O)\) precisely when
\[
  [T,B]=0,\qquad B\in\mathcal R(\mathcal O^\perp).
\]
This criterion is intrinsic to the represented net, but it is weaker than
membership in the originally assigned algebra \(\mathcal R(\mathcal O)\)
unless Haag duality has been established.  The dual net is therefore not a
decorative reformulation of locality.  It records the maximal bounded
operator algebra which the exterior observables cannot distinguish from a
local operator in \(\mathcal O\).

Two elementary consequences are used repeatedly later.  First, if
\(\mathcal O_1\subset\mathcal O_2\), then
\[
  \mathcal R^d(\mathcal O_1)
  =
  \mathcal R(\mathcal O_1^\perp)'
  \subset
  \mathcal R(\mathcal O_2^\perp)'
  =
  \mathcal R^d(\mathcal O_2),
\]
so the dual assignment is isotonic.  Second, if \(A\in\mathcal R(\mathcal O)\)
then locality gives
\[
  [A,B]=0,\qquad B\in\mathcal R(\mathcal O^\perp),
\]
hence \(A\in\mathcal R^d(\mathcal O)\).  Haag duality is exactly the assertion
that these two tests are equivalent for the chosen represented net.

\paragraph{Load-bearing use in localized sectors.}
Let \(\rho\) be a normal endomorphism of the represented quasilocal von
Neumann algebra which is localized in a causally complete region
\(\mathcal O_0\), in the sense that
\[
  \rho(B)=B,\qquad B\in\mathcal R(\mathcal O_0^\perp).
\]
Let \(\mathcal O_1\) be causally complete with
\(\mathcal O_0\subset\mathcal O_1\).  For
\(A\in\mathcal R(\mathcal O_1)\) and
\(B\in\mathcal R(\mathcal O_1^\perp)\), one has
\(\mathcal O_1^\perp\subset\mathcal O_0^\perp\), so \(\rho(B)=B\).  Since
\(A\) and \(B\) commute,
\[
  \rho(A)B
  =
  \rho(A)\rho(B)
  =
  \rho(AB)
  =
  \rho(BA)
  =
  \rho(B)\rho(A)
  =
  B\rho(A).
\]
Therefore
\[
  \rho(A)\in\mathcal R(\mathcal O_1^\perp)'
  =
  \mathcal R^d(\mathcal O_1).
\]
If Haag duality holds for \(\mathcal O_1\), this improves to
\[
  \rho(\mathcal R(\mathcal O_1))\subset\mathcal R(\mathcal O_1).
\]
This is the concrete algebraic step behind the DHR passage from a sector that
is indistinguishable from the vacuum outside \(\mathcal O_0\) to an
endomorphism of the local observable net.  Without the equality
\(\mathcal R=\mathcal R^d\), the same calculation gives an endomorphism into
the dual net; additional input is then required to identify that image with
the originally specified observable algebra.

The same point controls intertwiners.  If \(\rho\) and \(\sigma\) are
localized in \(\mathcal O\) and \(T\) satisfies
\[
  T\rho(A)=\sigma(A)T
\]
for all quasilocal \(A\), then for every
\(B\in\mathcal R(\mathcal O^\perp)\),
\[
  TB=T\rho(B)=\sigma(B)T=BT.
\]
Thus \(T\in\mathcal R^d(\mathcal O)\), and Haag duality is precisely what
places the intertwiner in the local algebra \(\mathcal R(\mathcal O)\).  The
tensor-category construction of bounded localized charges therefore uses
Haag duality as a structural hypothesis, not as a terminology choice.

\section{Point Fields as Coordinatizations}

If a Wightman field \(\Phi(f)\) is available, bounded functions of smeared
fields can generate local algebras.  Conversely, in many algebraic settings
point fields arise as operator-valued distributions affiliated with the local
algebras.  The invariant object is the local observable content:
\[
  \mathcal O\mapsto \Obs(\mathcal O)
  \quad\text{or}\quad
  \mathcal O\mapsto\mathcal R(\mathcal O).
\]
Fields are then coordinatizations of local observables, organized by symmetry,
scaling, and the operator product expansion when those structures are
available.

The passage from unbounded Wightman fields to a Haag--Kastler net is not a
formal consequence of the Wightman axioms in their bare domain formulation.
For real test functions the smeared fields are symmetric operators on a common
invariant domain, but this alone neither supplies self-adjoint closures nor
implies strong commutativity of the spectral measures of spacelike separated
closures.  The construction below records the standard route under the extra
closure and strong-locality hypotheses needed to make it a theorem.
The source traces for the two sides of this comparison are the Wightman and
Haag--Kastler frameworks
\cite{Wightman1956,StreaterWightman1964,HaagKastler1964}; the theorem below
isolates the extra functional-analytic hypotheses rather than folding them
into either named framework.

\begin{definition}[Affiliation of an unbounded operator]
\label{def:operator-affiliation-local-algebra}
For a self-adjoint operator \(T\) on \(\Hilb\), affiliation with a von
Neumann algebra \(\mathcal M\subset B(\Hilb)\) means that the spectral
projections \(E_T(\Delta)\) of \(T\) belong to \(\mathcal M\) for every Borel
set \(\Delta\subset\mathbb R\).  Equivalently, every unitary in the commutant
\(\mathcal M'\) preserves the graph of \(T\) and commutes with \(T\) on its
domain.
\end{definition}

Affiliation is the precise way in which an unbounded field can be localized
in a bounded local algebra.

The equivalence in Definition~\ref{def:operator-affiliation-local-algebra} is
often the practical test for locality of unbounded field coordinates, so we
spell it out.  Let \(E_T\) be the spectral measure of \(T\).  If
\(E_T(\Delta)\in\mathcal M\) for every Borel \(\Delta\), then every
\(B\in\mathcal M'\) commutes with every bounded Borel function of \(T\).  In
particular,
\[
  B(1+T^2)^{-1/2}=(1+T^2)^{-1/2}B,
  \qquad
  BT(1+T^2)^{-1/2}=T(1+T^2)^{-1/2}B .
\]
The first identity implies that \(B\operatorname{Dom}T\subset\operatorname{Dom}T\):
indeed
\[
  \operatorname{Dom}T=(1+T^2)^{-1/2}\Hilb .
\]
The second identity then gives \(BT\xi=TB\xi\) for
\(\xi\in\operatorname{Dom}T\), by applying it to
\(\xi=(1+T^2)^{-1/2}\eta\).  Thus commutant operators preserve the graph and
commute with \(T\) on the graph domain.

Conversely, suppose every unitary \(V\in\mathcal M'\) preserves
\(\operatorname{Dom}T\) and satisfies \(VT\xi=TV\xi\) for
\(\xi\in\operatorname{Dom}T\).  Then \(VT V^{-1}=T\) as self-adjoint
operators, so uniqueness in the spectral theorem gives
\[
  V E_T(\Delta)V^{-1}=E_T(\Delta)
\]
for every Borel \(\Delta\).  Hence \(E_T(\Delta)\) commutes with all unitaries
of \(\mathcal M'\).  The unitary group of a von Neumann algebra generates it
ultraweakly, and the bicommutant theorem gives
\[
  (\mathcal M')'=\mathcal M .
\]
Therefore \(E_T(\Delta)\in\mathcal M\).  This argument is the bounded
operator-algebra content behind the phrase ``the unbounded field is localized
in \(\mathcal M\)''; the difficult QFT input is proving that the closed field
operators in question have such spectral projections.

\begin{theorem}[Wightman fields generate a represented local net under strong closure hypotheses]
\label{thm:wightman-to-aqft-net}
Let
\[
  (\Hilb,\vac,U,\mathcal D,\{\widehat\Phi_\alpha\}_{\alpha\in I})
\]
be a Wightman field presentation on \(M\).  Enlarge the field-label space, if
needed, so that the smeared generators used below are Hermitian on
\(\mathcal D\).  Suppose that for every real Hermitian labelled test function
\(h\) the symmetric operator \(\widehat\Phi(h)\) is essentially self-adjoint on
\(\mathcal D\); write \(X(h)\) for its self-adjoint closure and
\[
  E_{h}(\Delta)
\]
for the spectral projection of \(X(h)\) associated with the Borel set
\(\Delta\subset\mathbb R\).  Define, for every region
\(\mathcal O\in\mathcal K(M)\),
\begin{equation}
  \mathcal R_\Phi(\mathcal O)
  :=
  \{E_h(\Delta)\mid
    \operatorname{supp}h\subset\mathcal O,\quad
    h=h^\dagger,\quad
    \Delta\subset\mathbb R\text{ Borel}\}'' .
  \label{eq:wightman-generated-net-from-spectral-projections}
\end{equation}
Equivalently, by the spectral theorem,
\begin{equation}
  \mathcal R_\Phi(\mathcal O)
  =
  \{\exp(\ii t X(h))\mid
    \operatorname{supp}h\subset\mathcal O,\quad
    h=h^\dagger,\quad
    t\in\mathbb R\}'' .
  \label{eq:wightman-generated-net-from-weyl-unitaries}
\end{equation}
Since \(t\) may be absorbed into \(h\), this is the standard displayed
construction
\[
  \mathcal R_\Phi(\mathcal O)
  =
  \{\exp(\ii X(h))\mid
    \operatorname{supp}h\subset\mathcal O,\quad h=h^\dagger\}'' .
\]
For every such \(h\), the closed operator \(X(h)\) is affiliated with
\(\mathcal R_\Phi(\mathcal O)\).  If
\(\mathcal O_1\subset\mathcal O_2\), the same \(X(h)\) is also affiliated with
\(\mathcal R_\Phi(\mathcal O_2)\).
Assume strong locality of these closures: if
\(\mathcal O_1\perp\mathcal O_2\), then every spectral projection entering
\(\mathcal R_\Phi(\mathcal O_1)\) commutes with every spectral projection
entering \(\mathcal R_\Phi(\mathcal O_2)\).  Then
\[
  \mathcal O\longmapsto\mathcal R_\Phi(\mathcal O)
\]
is an isotonic, Poincare-covariant represented local net of von Neumann
algebras on \(\Hilb\).  The vector state
\[
  A\longmapsto\langle\vac,A\vac\rangle
\]
is invariant under the represented Poincare action.
\end{theorem}

\begin{proof}
For a fixed \(h\), the spectral theorem identifies the von Neumann algebra
generated by the projection-valued measure \(E_h\) with the von Neumann
algebra generated by the one-parameter unitary group
\(\{\exp(\ii tX(h))\}_{t\in\mathbb R}\).  Indeed
\(\exp(\ii tX(h))=\int \exp(\ii t\lambda)\,\dd E_h(\lambda)\), so the
unitaries lie in
\(\{E_h(\Delta)\mid\Delta\subset\mathbb R\text{ Borel}\}''\); conversely, the
spectral projections are bounded Borel functions of \(X(h)\), hence lie in
the von Neumann algebra generated by the unitary group by the Borel functional
calculus for a self-adjoint operator.  This proves the equivalence of
\eqref{eq:wightman-generated-net-from-spectral-projections} and
\eqref{eq:wightman-generated-net-from-weyl-unitaries}, and also proves the
affiliation statement because all spectral projections of \(X(h)\) are
generators of \(\mathcal R_\Phi(\mathcal O)\).

If \(\mathcal O_1\subset\mathcal O_2\), then the test functions with support in
\(\mathcal O_1\) are among those with support in \(\mathcal O_2\), hence the
generating spectral projections for
\(\mathcal R_\Phi(\mathcal O_1)\) are contained in
\(\mathcal R_\Phi(\mathcal O_2)\).  This proves isotony.  Wightman covariance
gives
\[
  U(g)X(h)U(g)^{-1}=X(g h),
\]
where \(g h\) is the transformed labelled test function.  The spectral theorem
then gives
\[
  U(g)E_h(\Delta)U(g)^{-1}=E_{g h}(\Delta),
\]
so
\[
  U(g)\mathcal R_\Phi(\mathcal O)U(g)^{-1}
  =
  \mathcal R_\Phi(g\mathcal O).
\]
Strong locality is exactly the commutation of the spectral-projection
generators for spacelike separated regions; taking double commutants preserves
commutation of von Neumann algebras.  The vacuum invariance follows from
\(U(g)\vac=\vac\).
\end{proof}

The constructive checkpoint below records a useful sufficient mechanism for
the strong-locality hypothesis: Lemma~\ref{lem:analytic-vector-strong-locality}
shows how joint analytic-vector estimates upgrade domain-level Wightman
commutators to commutation of spectral projections.  This is an additional
functional-analytic input beyond the abstract Wightman distributional axioms.

There is an analogous point for additivity.  The formula
\eqref{eq:wightman-generated-net-from-spectral-projections} is isotonic by
inclusion of its generating spectral projections, but Haag--Kastler additivity
for a cover is not a formal consequence of writing a test function as a sum of
smaller test functions.  Suppose
\(\mathcal O\subset\bigcup_{j=1}^r\mathcal O_j\) and \(h\) is Hermitian with
\(\operatorname{supp}h\subset\mathcal O\).  A partition of unity gives
\[
  h=h_1+\cdots+h_r,\qquad
  \operatorname{supp}h_j\subset\mathcal O_j .
\]
On the Wightman domain one has
\(\widehat\Phi(h)=\sum_j\widehat\Phi(h_j)\), but the bounded local algebra
requires the spectral projections of the self-adjoint closure \(X(h)\).  A
sufficient additivity condition is therefore the closure/Trotter statement
\[
  X(h)=\overline{X(h_1)+\cdots+X(h_r)}
\]
on a common core, together with the strong Trotter product formula
\begin{equation}
  \exp(\ii tX(h))
  =
  \operatorname*{s-lim}_{n\to\infty}
  \left(
    \exp\!\left(\frac{\ii t}{n}X(h_1)\right)
    \cdots
    \exp\!\left(\frac{\ii t}{n}X(h_r)\right)
  \right)^n .
  \label{eq:wightman-net-trotter-additivity-bridge}
\end{equation}
The right side belongs to
\(\bigvee_j\mathcal R_\Phi(\mathcal O_j)\).  The spectral projections of
\(X(h)\), being bounded Borel functions of the unitary group
\(\{\exp(\ii tX(h))\}_{t\in\mathbb R}\), then belong to the same generated
von Neumann algebra.  Thus additivity of a Wightman-generated bounded net is
a closure and functional-calculus statement about the self-adjoint field
coordinates; it should not be imported from the distributional linearity of
the fields alone.  The finite check
\path{calculation-checks/wightman_net_bridge_checks.py} contains the
corresponding matrix-algebra shadow: the spectral projection of a sum of two
noncommuting local self-adjoint generators lies in the von Neumann algebra
generated by the two smaller algebras only after the sum has been specified as
an operator in that generated algebra.

\begin{remark}[Status of the Wightman-to-net direction]
The theorem is conditional.  Wightman commutativity is initially a statement
about commutators of unbounded smeared fields on a common dense domain; it
does not automatically imply commutativity of the spectral projections of
self-adjoint closures.  Thus the displayed Weyl-unitary formula should not be
read as saying that arbitrary Wightman fields automatically span a local
Haag--Kastler algebra.  In constructive scalar models obtained from Euclidean
measures, the route is model-specific: one proves the Schwinger functions,
OS reconstruction, Wightman properties, and the required local-net properties
with estimates particular to the construction.  That is a proof of the
Haag--Kastler axioms for those models, not a general implication from the
abstract Wightman axioms alone.
\end{remark}

\begin{definition}[Wightman-to-net comparison structure]
\label{def:wightman-to-net-comparison-structure}
Let
\((\Hilb,\vac,U,\mathcal D,\{\widehat\Phi_\alpha\}_{\alpha\in I})\)
be a Wightman field presentation.  A Wightman-to-net comparison structure for a
chosen region class \(\mathcal K(M)\) consists of the following additional
objects and hypotheses.
\begin{enumerate}[label=\textup{(\roman*)}]
\item For every \(\mathcal O\in\mathcal K(M)\), a real vector space
\[
  \mathcal H_\Phi(\mathcal O)
  \subset
  \operatorname{span}_{\mathbb R}
  \{\text{Hermitian labelled test functions supported in }\mathcal O\}
\]
such that
\(\mathcal O_1\subset\mathcal O_2\) implies
\(\mathcal H_\Phi(\mathcal O_1)\subset\mathcal H_\Phi(\mathcal O_2)\), and
\[
  U(g)\widehat\Phi(h)U(g)^{-1}=\widehat\Phi(gh),
  \qquad
  gh\in\mathcal H_\Phi(g\mathcal O)
\]
whenever \(h\in\mathcal H_\Phi(\mathcal O)\).
\item A self-adjoint closure \(X(h)=\overline{\widehat\Phi(h)}\) for every
\(h\in\mathcal H_\Phi(\mathcal O)\), with the covariance identity
\[
  U(g)X(h)U(g)^{-1}=X(gh)
\]
understood as equality of self-adjoint operators on their closed graph
domains.
\item Strong spacelike locality: if
\(\mathcal O_1\perp\mathcal O_2\),
\(h_i\in\mathcal H_\Phi(\mathcal O_i)\), and
\(E_{h_i}\) is the spectral measure of \(X(h_i)\), then
\begin{equation}
  E_{h_1}(\Delta_1)E_{h_2}(\Delta_2)
  =
  E_{h_2}(\Delta_2)E_{h_1}(\Delta_1)
  \label{eq:wightman-net-strong-locality-condition}
\end{equation}
for all Borel sets \(\Delta_1,\Delta_2\subset\mathbb R\).
\item An observable-subnet prescription.  If the fields include charged,
gauge-variant, or otherwise non-observable coordinates, the field-generated
net
\[
  \mathcal F_\Phi(\mathcal O)
  =
  \{E_h(\Delta)\mid h\in\mathcal H_\Phi(\mathcal O)\}''
\]
is distinct from the neutral observable net.  One must specify an observable
subnet, for example a fixed-point net
\[
  \mathcal R_\Phi(\mathcal O)=\mathcal F_\Phi(\mathcal O)^G
\]
for an implemented compact internal group \(G\), or a separately constructed
gauge-invariant subnet.
\end{enumerate}
\end{definition}

This comparison structure is deliberately stronger than the Wightman
distributional axioms.
The Wightman hierarchy determines the cyclic field representation and its
vacuum matrix elements.  It does not, by itself, decide which self-adjoint
closures have Borel functional calculi compatible with locality, nor does it
decide whether the field coordinates are observables.  Exponentials of
smeared Hermitian fields, when they exist, generate the displayed
field-generated net; in a theory with a nontrivial internal or gauge
structure this net may be larger than the neutral observable net.  Thus the
question ``do the exponentials of smeared Wightman fields span the local
algebra?'' has no framework-independent answer.  It becomes a theorem only
after the comparison structure identifies the field coordinates, their closures,
their strong locality, and the observable-subnet operation.

The finite algebra behind the last point is simple but important.  Let
\(\Gamma=\mathbb Z/2\mathbb Z\) act on a two-dimensional Hilbert space by the
parity unitary \(P=\operatorname{diag}(1,-1)\).  The self-adjoint operator
\(X=\begin{psmallmatrix}0&1\\1&0\end{psmallmatrix}\) is odd:
\(PXP^{-1}=-X\).  Its spectral projections generate bounded field
coordinates, but they are exchanged by \(\Gamma\) and therefore are not
observables of the fixed-point algebra.  The invariant subalgebra is the
diagonal algebra generated by \(1\) and \(P\), while adding the spectral
projections of \(X\) generates the full matrix algebra.  This toy model is
a bounded-algebra diagnostic; it supplies no spacetime localization,
domain, or spectrum theorem.  Its role is to check the field-net versus
observable-net distinction that the Wightman-to-net comparison structure must
retain in every genuine model.

\paragraph{Affiliation is not generation.}
The converse comparison is equally important.  Suppose a represented
observable net \(\mathcal R(\mathcal O)\) is already given, and suppose one
has a family of self-adjoint field coordinates \(T_i(f)\) affiliated with
\(\mathcal R(\mathcal O)\) whenever \(\operatorname{supp}f\subset\mathcal O\).
Affiliation means that all spectral projections of these coordinates are
bounded local observables.  It gives a canonical inclusion
\begin{equation}
  \mathcal R_{\rm coord}(\mathcal O)
  :=
  \{E_{T_i(f)}(\Delta)\mid
    \operatorname{supp}f\subset\mathcal O,
    \Delta\subset\mathbb R\text{ Borel}\}''
  \subset
  \mathcal R(\mathcal O),
  \label{eq:affiliated-coordinate-generated-subnet}
\end{equation}
not an equality.  Equality is a separate generation, or coordinatization
completeness, hypothesis:
\[
  \mathcal R_{\rm coord}(\mathcal O)=\mathcal R(\mathcal O)
  \qquad\text{for every region }\mathcal O
\]
in the region class under consideration.  This hypothesis may hold for a
chosen generating field system, may fail for a distinguished but incomplete
field such as a current or stress tensor, and may be inappropriate when the
physical observable net is a fixed-point or gauge-invariant subnet of a
larger field net.  Thus the statement that local fields are affiliated with
local algebras is weaker than the statement that their bounded functional
calculus generates all local observables.

The finite diagonal model makes the distinction visible without domains.  In
the algebra \(D_3\subset M_3(\mathbb C)\) of diagonal matrices, the
self-adjoint coordinate
\[
  T=\operatorname{diag}(0,0,1)
\]
is affiliated with \(D_3\), because its spectral projections
\(\operatorname{diag}(1,1,0)\) and \(\operatorname{diag}(0,0,1)\) lie in
\(D_3\).  They generate only the two-dimensional subalgebra
\(\mathbb C\operatorname{diag}(1,1,0)\oplus
\mathbb C\operatorname{diag}(0,0,1)\), not the full diagonal algebra.  A
second coordinate \(S=\operatorname{diag}(0,1,0)\) refines the spectral
partition and, together with \(T\), generates all three minimal diagonal
projections.  The finite check
\path{calculation-checks/wightman_net_bridge_checks.py} verifies exactly this
coarse-spectral-generation obstruction.  In QFT the analogous issue is not
finite-dimensional degeneracy but whether the chosen local fields separate
the local bounded observable content with the topology relevant to the
represented net.

\paragraph{Fixed points and additivity.}
Suppose the comparison structure has produced a represented field net
\(\mathcal F(\mathcal O)\), and suppose a compact internal group \(G\) is
implemented by strongly continuous unitaries \(V(g)\) such that
\[
  V(g)\vac=\vac,\qquad
  V(g)U(a,\Lambda)=U(a,\Lambda)V(g),
  \qquad
  V(g)\mathcal F(\mathcal O)V(g)^{-1}=\mathcal F(\mathcal O).
\]
The fixed-point candidate for the observable net is
\begin{equation}
  \mathcal R(\mathcal O)
  =
  \mathcal F(\mathcal O)^G
  =
  \{A\in\mathcal F(\mathcal O)\mid
    V(g)AV(g)^{-1}=A\text{ for all }g\in G\}.
  \label{eq:fixed-point-observable-subnet}
\end{equation}
The Haar integral gives a normal conditional expectation
\begin{equation}
  E_{\mathcal O}(A)
  =
  \int_G V(g)AV(g)^{-1}\,\dd g,
  \qquad
  A\in\mathcal F(\mathcal O),
  \label{eq:fixed-point-haar-expectation}
\end{equation}
where the integral is taken in the ultraweak topology.  It satisfies
\[
  E_{\mathcal O}^2=E_{\mathcal O},\qquad
  E_{\mathcal O}(B_1AB_2)=B_1E_{\mathcal O}(A)B_2,
  \quad B_1,B_2\in\mathcal R(\mathcal O),
\]
and its range is \(\mathcal R(\mathcal O)\).  Hence each
\(\mathcal R(\mathcal O)\) is a von Neumann algebra, isotony and Poincare
covariance are inherited from \(\mathcal F\), and locality follows because
\(\mathcal R(\mathcal O)\subset\mathcal F(\mathcal O)\).

The previous paragraph is only the bounded fixed-point mechanism.  It does
not by itself prove all Haag--Kastler properties for the observable net.
Additivity is the point where the distinction is visible.  For two regions
with field algebras carrying charged coordinates \(X_1\) and \(X_2\), a
neutral composite such as \(X_1X_2\) may be invariant under the diagonal
action of \(G\) on the algebra generated by the two field algebras, while
neither \(X_1\) nor \(X_2\) belongs to the separate fixed-point algebras.
Thus
\[
  \bigvee_i \mathcal F(\mathcal O_i)^G
  \subset
  \left(\bigvee_i\mathcal F(\mathcal O_i)\right)^G
\]
can be a proper inclusion.  An observable-net additivity theorem must explain
which neutral composites are generated locally, often using field-algebra
reconstruction, charged intertwiners, split inclusions, or a model-specific
constructive argument.  The formula
\eqref{eq:fixed-point-observable-subnet} supplies the bounded fixed-point
algebras; DHR/DR reconstruction or a constructive model proof supplies the
additional charged-sector and generation information needed for the full
observable Haag--Kastler axioms.

\paragraph{Constructive interacting checkpoint: massive \(P(\phi)_2\).}
The massive \(P(\phi)_2\) models developed constructively in
Chapter~\ref{ch:constructive-scalar-models-os-data} give a useful test of the
previous theorem because every word in the local-net construction has to be
earned from estimates.  Start with the infinite-volume Schwinger hierarchy of
Quoted Theorem~\ref{qthm:p-phi-two-cluster-os-output}, in a clustering
single-vacuum phase.  OS reconstruction gives a Hilbert space
\(\Hilb_P\), vacuum \(\Omega_P\), translation representation \(U_P\), and a
scalar Wightman field \(\Phi_P(f)\) on the reconstructed polynomial domain
\(\mathcal D_P\).  The polynomial domain is generated by positive-time
Euclidean fields modulo the OS null space; after analytic continuation it is
the common domain for finite products of smeared Lorentzian fields.

The associated bounded local field net is obtained from the spectral
projections of real smeared fields,
\begin{equation}
  \mathcal F_P(\mathcal O)
  =
  \{E_{\overline{\Phi_P(f)}}(\Delta)\mid
    f\in C_c^\infty(\mathcal O,\mathbb R),\
    \Delta\subset\mathbb R\text{ Borel}\}'' ,
  \label{eq:p-phi-two-constructive-local-field-net}
\end{equation}
where \(\overline{\Phi_P(f)}\) denotes the self-adjoint closure obtained in
the constructive Hilbert-space realization.  This displayed formula is a
definition of the net only after three analytic inputs have been supplied.
First, the Hamiltonian or Euclidean construction must prove essential
self-adjointness of the real smeared fields on a common invariant core.
Second, spacelike separated closures must commute strongly, meaning that their
spectral projections commute; the finite-cutoff equal-time tensor-factor
locality and the thermodynamic/ultraviolet limiting estimates are the model
mechanism for this statement.  Third, the limiting bounded functions must be
compatible with Poincare covariance and the time-slice dynamics.  With these
inputs, isotony follows from support inclusion, covariance from the
reconstructed \(U_P\), and locality from strong commutation of the spectral
projections.

\begin{lemma}[Strong-limit route from regulated locality to bounded locality]
\label{lem:strong-limit-regulated-locality}
Let \(\{A_\lambda\}_{\lambda\in\Lambda}\) and
\(\{B_\lambda\}_{\lambda\in\Lambda}\) be two nets of bounded operators on a
Hilbert space, indexed by a directed set.  Assume that
\[
  \sup_\lambda\|A_\lambda\|<\infty,\qquad
  \sup_\lambda\|B_\lambda\|<\infty,
  \qquad
  A_\lambda B_\lambda=B_\lambda A_\lambda ,
\]
and that \(A_\lambda\to A\), \(B_\lambda\to B\) strongly.  Then
\[
  AB=BA .
\]
Consequently, suppose a constructive regulator supplies self-adjoint
regulated local fields \(\Phi_\lambda(f)\) whose spectral projections
\(E_{\Phi_\lambda(f)}(\Delta)\) converge strongly to
\(E_{\overline{\Phi(f)}}(\Delta)\) for every interval \(\Delta\) in the
chosen spectral-window class.  If the finite-regulator projections commute
whenever \(\operatorname{supp}f\) and \(\operatorname{supp}g\) are spacelike
separated, then the limiting spectral projections commute for the same
windows.
\end{lemma}

\begin{proof}
For \(\xi\in\Hilb\),
\[
  \|(A_\lambda B_\lambda-AB)\xi\|
  \le
  \|A_\lambda\|\,\|(B_\lambda-B)\xi\|
  +
  \|(A_\lambda-A)B\xi\|.
\]
The first term tends to zero by strong convergence of \(B_\lambda\) and the
uniform bound on \(A_\lambda\).  The second tends to zero by strong
convergence of \(A_\lambda\), applied to the fixed vector \(B\xi\).  Hence
\(A_\lambda B_\lambda\to AB\) strongly.  The same argument gives
\(B_\lambda A_\lambda\to BA\) strongly.  Since
\(A_\lambda B_\lambda=B_\lambda A_\lambda\) for every \(\lambda\), their
strong limits are equal.

Apply this bounded statement to
\[
  A_\lambda=E_{\Phi_\lambda(f)}(\Delta),
  \qquad
  B_\lambda=E_{\Phi_\lambda(g)}(\Delta')
\]
for spacelike separated supports.  Spectral projections have norm at most
one, so the uniform boundedness hypothesis is automatic once the strong
spectral-window convergence is supplied.  The conclusion is the commutation
of the limiting spectral projections.  This proves bounded locality of the
local von Neumann algebras generated by those projections on the declared
spectral windows.
\end{proof}

This lemma is not a replacement for constructive estimates.  It identifies
the bounded convergence statement those estimates must produce.  A convergence
result only for vacuum expectation values of unbounded smeared fields, or only
for formal commutators on a polynomial domain, does not by itself imply
commutation of the spectral projections that define
\(\mathcal F_P(\mathcal O)\).  A model proof may use the strong-limit route
above, the analytic-vector route of
Lemma~\ref{lem:analytic-vector-strong-locality}, or another functional
calculus argument; it must supply one of them before the Wightman polynomial
field presentation has become a bounded Haag--Kastler net.

If the polynomial \(P\) has a finite internal symmetry group \(\Gamma\), for
example \(\Gamma=\mathbb Z_2\) when \(P\) is even and the chosen phase is
\(\mathbb Z_2\)-invariant, the observable net is the fixed-point net
\begin{equation}
  \mathcal R_P(\mathcal O)
  =
  \mathcal F_P(\mathcal O)^\Gamma .
  \label{eq:p-phi-two-constructive-observable-net}
\end{equation}
Thus the same constructive model distinguishes the field net, the observable
net, and possible sector questions.  In broken phases, kink sectors are
localized by asymptotic boundary conditions and therefore sit at the boundary
between ordinary local endomorphism sectors and more global charge data.
Massive
clustering gives a spectral gap through the argument in
Chapter~\ref{ch:constructive-scalar-models-os-data}; nuclearity, split
inclusions, Haag duality, and DHR sector reconstruction require separate
model estimates.  The example is therefore an interacting AQFT object: it
supplies concrete local von Neumann algebras and also identifies which
algebraic properties require additional theorems beyond the OS axioms.

\paragraph{Gauge-theory checkpoint: the local Schwinger-model observable net.}
The massless Schwinger model of Chapter~\ref{ch:schwinger-model} supplies a
different test: the starting variables are a gauge potential and a charged
Dirac field, but the physical local algebra is generated by gauge-invariant
fields.  Let \(e>0\) be the dynamical charge, let
\(m_{\rm Sch}^2=e^2/\pi\), and fix a local electric-flux branch and a value of
the theta parameter as in Lemma~\ref{lem:schwinger-exact-bosonized-gauge-sector}.
The exact local bosonization result gives the operator identity
\begin{equation}
  E
  =
  -\frac{e}{\sqrt\pi}
  \left(\varphi+\frac{\theta}{2\sqrt\pi}\right),
  \qquad
  (\partial^\mu\partial_\mu-m_{\rm Sch}^2)E=0,
  \label{eq:schwinger-aqft-electric-field-net-generator}
\end{equation}
on the gauge-invariant local sector.  The additive constant in
\eqref{eq:schwinger-aqft-electric-field-net-generator} shifts smeared fields
by a scalar multiple of the identity and therefore does not change the von
Neumann algebras generated by their spectral projections.  Consequently the
local electric-field net is the massive scalar Weyl net of the benchmark
model above at mass \(m_{\rm Sch}\):
\begin{equation}
  \mathcal R_{\rm Sch}(\mathcal O)
  =
  \left\{
    E_{\overline{E(f)}}(\Delta)
    \mid f\in C_c^\infty(\mathcal O,\mathbb R),\
       \Delta\subset\mathbb R\text{ Borel}
  \right\}''
  \cong
  \Obs_{m_{\rm Sch}}(\mathcal O).
  \label{eq:schwinger-aqft-observable-net}
\end{equation}
This is an equality of represented local nets after the local bosonization
map has been fixed; it is not a statement that the gauge representative
\(a_\mu\) or the charged fermion \(\psi\) is an observable field.
Gauge-invariant currents are derivatives of the same scalar field, and hence
lie in the local net generated by \(E\), up to the already fixed contact-term
convention of the current algebra.  Wilson lines and line-dressed charged
fields are different data: they are external probes or sector-changing
operators with endpoint, path, or asymptotic flux information.  They are
therefore not elements of \(\mathcal R_{\rm Sch}(\mathcal O)\) for a bounded
region unless an additional construction supplies a gauge-invariant local
composite.

This example isolates the AQFT bridge in a solvable gauge theory.  The
Haag--Kastler properties of \(\mathcal R_{\rm Sch}\) follow from the massive
scalar Weyl net once the Schwinger-model derivation has proved the local
bosonization identity and the Gauss-law constraint on physical states.
External fractional probes, compact zero modes on a circle, and flux sectors
are not defects of the local net; they are additional sector data attached to
the same observable algebra.  Thus the example separates three notions that
are often conflated: the gauge representative used in a Lagrangian
calculation, the gauge-invariant local observable net, and the charged or
line-dressed sectors on which nonlocal probes act.

\begin{lemma}[Analytic-vector route to strong locality]
\label{lem:analytic-vector-strong-locality}
Let \(X\) and \(Y\) be symmetric operators on a Hilbert space \(\Hilb\), with
a common dense invariant domain \(\mathcal D\).  Assume:
\begin{enumerate}[label=\textup{(\roman*)}]
\item \(X\) and \(Y\) are essentially self-adjoint on \(\mathcal D\);
\item every vector in \(\mathcal D\) is jointly entire analytic for the
ordered pair \((X,Y)\), meaning that, for every \(\xi\in\mathcal D\) and
every \(r,s\geq0\),
\[
  \sum_{m,n\geq0}\frac{r^m s^n}{m!\,n!}\,\|X^mY^n\xi\|<\infty ;
\]
\item \([X,Y]\xi=0\) for every \(\xi\in\mathcal D\).
\end{enumerate}
Then the self-adjoint closures \(\overline X\) and \(\overline Y\) strongly
commute, equivalently all spectral projections of \(\overline X\) commute with
all spectral projections of \(\overline Y\).
\end{lemma}

\begin{proof}
Because \(\mathcal D\) is invariant under \(X\) and \(Y\), the commutator
assumption implies \(X^mY^n\xi=Y^nX^m\xi\) for all \(m,n\geq 0\) and
\(\xi\in\mathcal D\), by induction.  The joint analyticity assumption implies
separate analyticity for \(X\) and \(Y\), and also gives absolute convergence
of the double series below.  Let \(U(t)=\exp(\ii t\overline X)\) and
\(V(s)=\exp(\ii s\overline Y)\).  For \(\xi\in\mathcal D\), the joint
analytic-vector hypothesis gives norm-convergent power series
\[
  U(t)\xi=\sum_{m\geq0}\frac{(\ii t)^m}{m!}X^m\xi,\qquad
  V(s)\xi=\sum_{n\geq0}\frac{(\ii s)^n}{n!}Y^n\xi .
\]
The Cauchy product is legitimate because
\[
  \sum_{m,n\geq0}
  \frac{|t|^m|s|^n}{m!\,n!}\,\|X^mY^n\xi\|<\infty ,
\]
and gives
\[
  U(t)V(s)\xi
  =
  \sum_{m,n\geq0}\frac{(\ii t)^m(\ii s)^n}{m!\,n!}X^mY^n\xi
  =
  \sum_{m,n\geq0}\frac{(\ii t)^m(\ii s)^n}{m!\,n!}Y^nX^m\xi
  =
  V(s)U(t)\xi .
\]
Since \(\mathcal D\) is dense and \(U(t),V(s)\) are bounded unitaries, the
equality holds on all of \(\Hilb\).  The spectral theorem then implies
commutation of the projection-valued measures of \(\overline X\) and
\(\overline Y\).
\end{proof}

For the constructive net
\(\mathcal O\mapsto\mathcal F_P(\mathcal O)\), this lemma states exactly
which analytic estimate must accompany Wightman locality.  If
\(f,g\in C_c^\infty(M,\mathbb R)\) have spacelike separated supports,
Wightman locality gives
\[
  [\Phi_P(f),\Phi_P(g)]\xi=0,\qquad \xi\in\mathcal D_P .
\]
The bounded local algebras in
\eqref{eq:p-phi-two-constructive-local-field-net} are local only after the
specific construction supplies a common core on which the smeared fields are
essentially self-adjoint and analytic-vector estimates hold.  In Hamiltonian
constructions these estimates usually come from comparison of
smeared fields with powers of the finite-volume Hamiltonian, followed by
uniform bounds in the infinite-volume and ultraviolet limits.  In Euclidean
constructions the same information is encoded in hypercontractive or
Nelson-type bounds for polynomial fields on the OS Hilbert space.  The
resulting strong commutation is a bounded-algebra statement; it is stronger
than the formal vanishing of a commutator on the polynomial domain.

The structural information obtained from the model should therefore be read
in the following order.  The OS hierarchy first gives the Hilbert space,
vacuum, translation representation, Wightman distributions, and polynomial
field domain.  Essential self-adjointness and analytic-vector estimates then
turn real smeared fields into spectral projection generators.  Strong
commutation of these generators gives locality of
\(\mathcal F_P\).  Internal fixed points give the observable net
\(\mathcal R_P=\mathcal F_P^\Gamma\).  Further AQFT properties are separate:
nuclearity is an estimate on
\[
  \Theta_{\beta,\mathcal O}\colon
  \mathcal F_P(\mathcal O)\to\Hilb_P,\qquad
  A\longmapsto \exp(-\beta H_P)A\Omega_P ,
\]
split inclusions follow only after such maps are trace-class or nuclear with
the appropriate local bounds, Haag duality is a statement about equality of a
local algebra with a commutant of an exterior algebra, and DHR reconstruction
requires a tensor category of transportable localized endomorphisms.  None of
these conclusions is contained in the symbol \(P(\phi)_2\) itself.  The
constructive example is valuable precisely because each assertion can be
attached to a named estimate or representation-theoretic construction.

This separation is also the right language for phase structure.  In an
unbroken finite-symmetry phase the fixed-point observable net has sectors
created by localized charged fields in the larger field net, when such fields
are available.  In a broken phase the natural kink Hilbert spaces are built
with different asymptotic boundary conditions at spatial infinity.  A kink
cannot be treated as an ordinary DHR charge in the vacuum representation
unless one proves an additional localization/transportability statement that
turns the boundary-condition sector into an endomorphism localized in bounded
regions.  Thus the example exhibits both the power and the limitations of the
standard local-endomorphism formulation of superselection theory.

\begin{remark}[Field coordinates as a Wightman presentation]
\label{rem:aqft-to-wightman-fields}
Let
\[
  \mathcal O\longmapsto\mathcal R(\mathcal O)
\]
be a Poincare-covariant represented local net on a Hilbert space \(\Hilb\),
with unitary implementation \(U\) and invariant unit vector \(\Omega\).  Assume
the translation spectrum of \(U\) is contained in \(\overline V_+\).  Suppose
there are a dense subspace \(\mathcal D\subset\Hilb\), a label set \(I\), and
operators \(\widehat\Phi_\alpha(f)\) such that:
\begin{enumerate}[label=\textup{(\roman*)}]
\item \(\Omega\in\mathcal D\), \(U(g)\mathcal D\subset\mathcal D\), and
\(\widehat\Phi_\alpha(f)\mathcal D\subset\mathcal D\);
\item for each \(\alpha\), the map
\(f\mapsto\widehat\Phi_\alpha(f)\) is an operator-valued tempered
distribution on the common domain \(\mathcal D\);
\item if \(\operatorname{supp}f\subset\mathcal O\), then
\(\widehat\Phi_\alpha(f)\) is affiliated with \(\mathcal R(\mathcal O)\);
\item the fields satisfy the Wightman covariance and adjunction identities on
\(\mathcal D\);
\item fields affiliated with spacelike separated regions have vanishing graded
commutator on \(\mathcal D\);
\item the polynomial field orbit of \(\Omega\) is dense in \(\Hilb\).
\end{enumerate}
Then
\[
  (\Hilb,\Omega,U,\mathcal D,\{\widehat\Phi_\alpha\}_{\alpha\in I})
\]
is a Wightman field presentation, and its vacuum distributions are the matrix
elements
\[
  W_{\alpha_1\cdots\alpha_n}(f_1,\ldots,f_n)
  =
  \langle\Omega,
  \widehat\Phi_{\alpha_1}(f_1)\cdots
  \widehat\Phi_{\alpha_n}(f_n)\Omega\rangle .
\]
Indeed, the Hilbert space, invariant vacuum vector, unitary Poincare
representation, dense common domain, and operator-valued tempered
distributions are the data just listed.  The spectral condition is the assumed
joint spectrum condition for translations; covariance, adjunction, locality,
and cyclicity are items (iv), (v), and (vi).  Continuity of the vacuum matrix
elements is part of the operator-valued tempered-distribution hypothesis on
the common invariant domain.  Thus the listed AQFT-plus-field-coordinate data
are exactly a Wightman field presentation in the sense used in
Chapter~\ref{ch:volume-iv-wightman-fields-reconstruction}.
\end{remark}

This relation is important for gauge theories.  The algebraic observable net
can be defined from gauge-invariant observables without choosing gauge
potentials as physical operators.  If charged fields are introduced, they
belong to additional data: an enlarged field algebra, charged
representations, or sector endomorphisms.  They are not elements of the
neutral observable algebra unless the construction explicitly identifies them
as observables.

\section{Reeh--Schlieder and Local Cyclicity}

\begin{definition}[Weak additivity in a vacuum representation]
\label{def:weak-additivity-vacuum-net}
Let \(\vac\) be the vacuum vector in a Poincare-covariant representation, and
let
\[
  \mathcal R_{\rm glob}
  :=
  \bigvee_{\mathcal O\in\mathcal K(M)}\mathcal R(\mathcal O).
\]
Here \(\bigvee\) denotes the von Neumann algebra generated by the displayed
family.  The represented net is weakly additive when, for every nonempty open
region \(\mathcal U\) with nontrivial causal completion,
\[
  \bigvee_{a\in\mathbb R^D}\mathcal R(\mathcal U+a)
  =
  \mathcal R_{\rm glob}.
\]
\end{definition}

Weak additivity is the translated-generation condition used in the proof
below.  Finite-cover additivity identifies the algebra of one region with the
algebra generated by a cover of that same region; weak additivity is a global
condition on translates of one fixed local algebra.

\begin{lemma}[Positive-energy tube uniqueness for translation matrix coefficients]
\label{lem:positive-energy-tube-uniqueness}
Let \(M=\mathbb R^{1,D-1}\) carry the mostly-plus metric, and let
\(\mu\) be a finite complex Borel measure on momentum space with
\[
  \operatorname{supp}\mu\subset\overline V_+ .
\]
Define, for \(a\in M\),
\[
  F_0(a)=\int_{\overline V_+}\ee^{\ii a\cdot p}\,\dd\mu(p).
\]
For \(b\in V_+\), set
\[
  F(a-\ii b)
  =
  \int_{\overline V_+}\ee^{\ii(a-\ii b)\cdot p}\,\dd\mu(p).
\]
Then \(F\) is holomorphic in the tube \(M-\ii V_+\) and has continuous
boundary value \(F_0\) as \(b\downarrow0\).  If \(F_0\) vanishes on a
nonempty open subset of the real edge \(M\), then \(F_0\) vanishes on all of
\(M\).
\end{lemma}

\begin{proof}
For each compact subcone \(K\Subset V_+\) there is a constant \(c_K>0\) such
that
\[
  b\cdot p\le -c_K |p|,
  \qquad b\in K,\quad p\in\overline V_+ .
\]
Here \(|p|\) is any fixed Euclidean norm after choosing inertial coordinates;
changing this norm only changes \(c_K\).
Consequently
\[
  \left|\ee^{\ii(a-\ii b)\cdot p}\right|
  =
  \ee^{b\cdot p}
  \le
  \ee^{-c_K|p|},
\]
and the same exponential bound, multiplied by any fixed polynomial in \(p\),
dominates the derivatives obtained by differentiating under the integral.
This proves holomorphy in \(M-\ii V_+\).  Since \(\mu\) is finite and
\(\ee^{b\cdot p}\le1\), dominated convergence gives the continuous boundary
value \(F_0\) as \(b\downarrow0\).

Let \(O\subset M\) be a nonempty open set on which \(F_0=0\).  On the
opposite tube \(M+\ii V_+\) take the holomorphic function \(0\).  The boundary
values of \(F\) and \(0\) agree on \(O\).  Applying the distributional
edge-of-the-wedge theorem,
Theorem~\ref{thm:distributional-edge-of-the-wedge}, with cones
\(-V_+\) and \(V_+\), gives a connected complex neighborhood \(\mathcal N\)
of a smaller open ball \(B\Subset O\) and a holomorphic function on
\(\mathcal N\) agreeing with \(F\) on
\(\mathcal N\cap(M-\ii V_+)\) and with \(0\) on
\(\mathcal N\cap(M+\ii V_+)\).  The latter set is open in \(\mathcal N\), so
the identity theorem makes the glued holomorphic function identically zero on
\(\mathcal N\).  Hence \(F\) vanishes on the nonempty open set
\(\mathcal N\cap(M-\ii V_+)\).  Since the tube \(M-\ii V_+\) is connected,
the identity theorem again gives \(F=0\) on the whole tube.  Taking the
continuous boundary value proves \(F_0(a)=0\) for every real translation
\(a\).
\end{proof}

\noindent
The source trace for Reeh--Schlieder local cyclicity is
\cite{ReehSchlieder1961,StreaterWightman1964}.  The proof is included in the
represented-net form because the spectral condition and weak additivity are
the inputs needed later in modular and scattering arguments.

\begin{theorem}[Reeh--Schlieder theorem for a weakly additive vacuum net]
\label{thm:reeh-schlieder-weakly-additive-net}
Let \(\mathcal R\) be a Poincare-covariant represented local net on
\(\Hilb\).  Assume:
\begin{enumerate}[label=\textup{(\roman*)}]
\item the vacuum vector \(\vac\) is translation invariant;
\item the joint translation spectrum lies in \(\overline V_+\);
\item \(\vac\) is cyclic for \(\mathcal R_{\rm glob}\);
\item the represented net is weakly additive in the sense of
Definition~\ref{def:weak-additivity-vacuum-net}.
\end{enumerate}
Then for every nonempty open region \(\mathcal O\) containing the closure of
some nonempty open \(\mathcal U\in\mathcal K(M)\),
\[
  \overline{\mathcal R(\mathcal O)\vac}=\Hilb .
\]
\end{theorem}

\begin{proof}
Let \(\psi\) be orthogonal to
\(\mathcal R(\mathcal O)\vac\).  Choose a nonempty open region
\(\mathcal U\) whose closure is contained in \(\mathcal O\), and choose
\(A\in\mathcal R(\mathcal U)\).  For translations \(a\) in a neighborhood of
\(0\), the inclusion \(\mathcal U+a\subset\mathcal O\) and covariance give
\(U(a)AU(a)^{-1}\in\mathcal R(\mathcal O)\), hence
\[
  \langle \psi,\,U(a)A\vac\rangle=0
\]
there.  Let \(E\) be the joint spectral measure of the translation
generators, so that \(U(a)=\int e^{ia\cdot p}\,dE(p)\) and
\(\operatorname{supp}E\subset\overline V_+\).  The finite complex measure
\[
  \mu_{\psi,A}(\Delta)
  :=
  \langle\psi,E(\Delta)A\vac\rangle
\]
gives, for real \(a\),
\[
  F_A(a):=\langle\psi,U(a)A\vac\rangle
  =
  \int_{\overline V_+} e^{ia\cdot p}\,d\mu_{\psi,A}(p).
\]
For \(z=a-\ii b\) with \(b\in V_+\), the same spectral integral
\[
  F_A(z)
  =
  \int_{\overline V_+} e^{iz\cdot p}\,d\mu_{\psi,A}(p)
\]
is holomorphic in \(z\) by
Lemma~\ref{lem:positive-energy-tube-uniqueness}.  The same lemma applies
because \(F_A(a)\) vanishes for \(a\) in a neighborhood of \(0\).  Hence
\[
  \langle \psi,U(a)A\vac\rangle=0
\]
for every real translation \(a\).  Since \(A\) was arbitrary in
\(\mathcal R(\mathcal U)\), weak additivity says that \(\psi\) is orthogonal
to \(\mathcal R_{\rm glob}\vac\).  Global cyclicity then gives \(\psi=0\).
The orthogonal complement of \(\mathcal R(\mathcal O)\vac\) is zero, which
proves the density.
\end{proof}

\begin{corollary}[Separatingness from Reeh--Schlieder and locality]
\label{cor:reeh-schlieder-separatingness}
Under the hypotheses of
Theorem~\ref{thm:reeh-schlieder-weakly-additive-net}, let
\(\mathcal O\) be a nonempty open region to which the theorem applies.
If \(B\in\mathcal B(\Hilb)\) commutes with \(\mathcal R(\mathcal O)\) and
\(B\vac=0\), then \(B=0\).  Consequently \(\vac\) is separating for every
von Neumann algebra \(\mathcal M\subset\mathcal R(\mathcal O)'\).

In particular, let \(W\) be a wedge and define \(\mathcal R(W)\) as the von
Neumann algebra generated by double-cone algebras contained in \(W\).  If the
causal complement \(W'\) contains a nonempty open region \(\mathcal O\) to
which Theorem~\ref{thm:reeh-schlieder-weakly-additive-net} applies, then
\(\vac\) is separating for \(\mathcal R(W)\).
\end{corollary}

\begin{proof}
Let \(B\) commute with \(\mathcal R(\mathcal O)\) and satisfy \(B\vac=0\).
For every \(A\in\mathcal R(\mathcal O)\),
\[
  B A\vac=A B\vac=0 .
\]
Theorem~\ref{thm:reeh-schlieder-weakly-additive-net} says that
\(\mathcal R(\mathcal O)\vac\) is dense in \(\Hilb\).  Since \(B\) is
bounded, \(B\) vanishes on all of \(\Hilb\).  If
\(\mathcal M\subset\mathcal R(\mathcal O)'\), the same argument applies to
each \(B\in\mathcal M\) with \(B\vac=0\), which is precisely separatingness
of \(\vac\) for \(\mathcal M\).

For the wedge statement, locality gives
\(\mathcal R(W)\subset\mathcal R(\mathcal O)'\) whenever
\(\mathcal O\subset W'\).  The preceding paragraph then applies with
\(\mathcal M=\mathcal R(W)\).  No Haag duality is used here; Haag duality
would identify a larger commutant with an algebra assigned to the causal
complement, whereas the separatingness conclusion only needs the inclusion
forced by locality.
\end{proof}

The theorem and its separatingness corollary are used as inputs for modular
theory, superselection analysis, and scattering theory in algebraic form.

Figure~\ref{fig:reeh-schlieder-cyclicity} emphasizes the strong content of the
theorem.  Operators localized in a fixed small region \(\mathcal O\) generate,
from the vacuum alone, a set of vectors dense in the whole vacuum Hilbert
space.  The conclusion is a Hilbert-space density statement controlled by
positive-energy analytic continuation: weak additivity and analyticity leave
no nonzero vector orthogonal to all vectors \(A\vac\) with
\(A\in\mathcal R(\mathcal O)\).

\begin{figure}[H]
  \centering
  \resizebox{0.95\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      arrow/.style={-{Latex[length=2mm]},line width=0.85pt},
      lab/.style={fill=white,inner sep=1.6pt,align=center},
      cone/.style={line width=0.85pt}
    }
    \node[font=\scriptsize\bfseries,align=center] at (-4.85,2.55)
      {real translations\\of the local region};
    \node[font=\scriptsize\bfseries,align=center] at (2.45,2.55)
      {tube analyticity\\and boundary uniqueness};

    \draw[->] (-7.05,-1.45)--(-2.35,-1.45);
    \node[lab] at (-2.45,-1.70) {\(x^1\)};
    \draw[->] (-6.45,-1.78)--(-6.45,1.92) node[above] {\(x^0\)};
    \draw[cone,blue!70!black,fill=blue!6]
      (-6.45,1.25)--(-5.28,0.05)--(-6.45,-1.15)--(-7.62,0.05)--cycle;
    \node[lab,text=blue!70!black] at (-7.55,1.22) {\(\mathcal O\)};
    \draw[cone,red!72!black,fill=red!12]
      (-6.45,0.56)--(-5.96,0.05)--(-6.45,-0.46)--(-6.94,0.05)--cycle;
    \node[lab,text=red!72!black] at (-6.13,0.83) {\(\mathcal U\Subset\mathcal O\)};
    \foreach \x/\opac in {-6.02/36,-5.60/28} {
      \draw[cone,red!\opac!black,fill=red!\opac!white]
        (\x,0.56)--(\x+0.49,0.05)--(\x,-0.46)--(\x-0.49,0.05)--cycle;
    }
    \node[lab] at (-5.78,-0.92) {small \(a\)};
    \foreach \x/\y in {-4.08/0.35,-3.32/-0.35,-2.65/0.52} {
      \draw[cone,blue!30!black,fill=blue!5]
        (\x,\y+0.44)--(\x+0.42,\y)--(\x,\y-0.44)--(\x-0.42,\y)--cycle;
    }
    \draw[arrow,blue!65!black] (-5.45,1.02) .. controls (-4.65,1.72) and (-3.55,1.58) .. (-2.80,0.98)
      node[pos=0.54,above,lab] {all real translates};
    \node[lab] at (-3.35,-1.03)
      {weak additivity:\\all translates generate\\\(\mathcal R_{\rm glob}\)};

    \draw[->] (-0.45,-1.45)--(5.20,-1.45) node[right] {\(\operatorname{Re} a\)};
    \draw[->] (0.05,-1.80)--(0.05,1.95) node[above] {tube depth \(\eta\in V_+\)};
    \fill[orange!12] (-0.22,-1.22)--(4.65,-1.22)--(3.43,1.55)--(0.95,1.55)--cycle;
    \draw[orange!75!black,line width=0.9pt]
      (-0.22,-1.22)--(4.65,-1.22)--(3.43,1.55)--(0.95,1.55)--cycle;
    \node[lab,text=orange!80!black] at (3.83,1.18)
      {\(\mathbb R^D-\ii V_+\)};
    \draw[red!75!black,line width=1.25pt] (0.55,-1.45)--(1.70,-1.45);
    \node[lab,text=red!75!black] at (1.13,-1.08)
      {boundary zeros\\near \(0\)};
    \draw[arrow,orange!80!black] (1.78,-1.10) .. controls (2.24,-0.36) and (2.47,0.62) .. (2.25,1.30);
    \draw[arrow,orange!80!black] (2.56,0.95) .. controls (3.15,0.50) and (3.55,-0.22) .. (3.72,-1.08);
    \node[lab] at (2.24,0.17)
      {\(F_A(z)=\int\ee^{\ii z\cdot p}\dd\mu_{\psi,A}(p)\)\\is analytic};
    \node[lab] at (3.58,-1.03)
      {\(F_A(a)=0\)\\for all real \(a\)};

    \node[align=center,fill=green!5,draw=green!45!black,line width=0.75pt,
      rounded corners=3pt,inner xsep=5pt,inner ysep=4pt] at (-0.75,-2.95)
      {\(\psi\perp\mathcal R(\mathcal O)\vac\)
      \(\Rightarrow\)
      \(F_A(a)=0\) near \(0\)
      \(\Rightarrow\)
      \(F_A(a)=0\) for all \(a\)
      \(\Rightarrow\)
      \(\psi\perp\mathcal R_{\rm glob}\vac\)
      \(\Rightarrow\)
      \(\psi=0\)};
  \end{tikzpicture}
  }
  \caption{Reeh--Schlieder local cyclicity.  The geometric input is the
  existence of \(\mathcal U\Subset\mathcal O\), so that small real
  translations of \(\mathcal U\) remain inside \(\mathcal O\).  If
  \(\psi\perp\mathcal R(\mathcal O)\vac\), then
  \(F_A(a)=\langle\psi,U(a)A\vac\rangle\) vanishes on a real neighborhood of
  \(0\) for every \(A\in\mathcal R(\mathcal U)\).  The spectral condition
  analytically continues \(F_A\) to the forward spectral tube
  \(M-\ii V_+\); the boundary-value
  uniqueness lemma for such tube-holomorphic functions propagates this
  vanishing to all real translations.  Weak additivity then turns the
  translated local algebras into \(\mathcal R_{\rm glob}\), and global
  cyclicity forces \(\psi=0\).}
  \label{fig:reeh-schlieder-cyclicity}
\end{figure}

\section{Locally Covariant Form}

\begin{definition}[Locally covariant \(C^*\)-algebraic QFT functor]
\label{def:locally-covariant-qft-functor}
Let \(\mathbf{Loc}\) be the category whose objects are oriented,
time-oriented globally hyperbolic Lorentzian spacetimes \(X\), and whose
morphisms \(\chi:X\to Y\) are isometric embeddings preserving orientation and
time orientation, with causally convex image.  Let \(\mathbf{C^*Alg}_{\rm inj}\)
be the category of unital \(C^*\)-algebras and injective unital
\(*\)-homomorphisms.  A locally covariant \(C^*\)-algebraic QFT is a functor
\[
  \mathfrak A:\mathbf{Loc}\longrightarrow\mathbf{C^*Alg}_{\rm inj}.
\]
For a morphism \(\chi:X\to Y\), write
\[
  \alpha_\chi:=\mathfrak A(\chi):\mathfrak A(X)\to\mathfrak A(Y).
\]
The functor satisfies the time-slice property when \(\alpha_\chi\) is an
isomorphism for every Cauchy morphism, meaning every \(\chi\) whose image
contains a Cauchy surface of \(Y\).  It satisfies Einstein causality when, for
two morphisms \(\chi_i:X_i\to Y\) with causally disjoint images,
\[
  [\alpha_{\chi_1}(A_1),\alpha_{\chi_2}(A_2)]=0,
  \qquad
  A_i\in\mathfrak A(X_i).
\]
\end{definition}

This formulation preserves the local content of QFT in spacetimes where
symmetry selects no distinguished global vacuum.  On Minkowski spacetime,
restricting \(\mathfrak A\) to embeddings of suitable regions into \(M\)
recovers a Haag--Kastler-type net.

For the present monograph, the algebraic framework supplies a precise language
for local observables, states, sectors, and structural locality theorems.  Its
connection with path integrals, perturbative expansions, and point-field
descriptions is a comparison problem to be tracked explicitly case by case.
