\chapter{Osterwalder--Schrader Reconstruction}
\label{ch:volume-iv-osterwalder-schrader-reconstruction}
\ConventionsMixed

\label{ch:os-reconstruction}

\section{Euclidean Correlation Data}

Let \(E=\mathbb R^D\) be Euclidean space with coordinates
\[
  x=(\tau,\mathbf x),\qquad \mathbf x\in\mathbb R^{D-1}.
\]
Let \(V\) be the complex vector space of field labels.  For scalar fields
\(V=\mathbb C\).  For tensor, spinor, or internal multiplets, \(V\) carries the
corresponding Euclidean rotation and internal representations.  An \(n\)-point
Euclidean test insertion is an element of
\[
  \mathcal S(E^n;V^{\otimes n}),
\]
or of \(C_c^\infty(E^n;V^{\otimes n})\) if compactly supported test functions
are the chosen test class.  The test class is part of the data because it
determines the topology in which the Schwinger functions are distributions.
Let
\[
  \Delta_n
  =
  \bigcup_{1\le i<j\le n}\{(x_1,\ldots,x_n)\in E^n\mid x_i=x_j\}
\]
be the union of partial diagonals.  The corrected OS reconstruction theorem is
most honestly stated on the zero-diagonal Schwartz space
\[
  \mathcal S_\circ(E^n;V^{\otimes n})
  =
  \{f\in\mathcal S(E^n;V^{\otimes n})
    \mid \partial^\alpha f|_{\Delta_n}=0\text{ for every multi-index }\alpha\}.
\]
Pairing with \(\mathcal S_\circ\) treats the Schwinger functions as
distributions away from coincident points.  A choice of contact-term extension
to the full Schwartz space is extra structure and is not part of the OS
reconstruction input used here.

\begin{definition}[Schwinger hierarchy]
A Schwinger hierarchy with field-label space \(V\) is a family
\[
  S_n\in \mathcal S_\circ'(E^n;(V^*)^{\otimes n}),
  \qquad n\ge0,
\]
with \(S_0=1\).  The value \(S_n(f_n)\) denotes the distributional pairing of
\(S_n\) with the \(V^{\otimes n}\)-valued test function \(f_n\).  In
component notation this is written as
\[
  S_{\alpha_1\cdots\alpha_n}(x_1,\ldots,x_n),
  \qquad \alpha_j\in I,
\]
where \(I\) is a basis of labels for \(V\).  The component notation is a
kernel notation for the distributional pairing; the pairing with test
functions is the primary object.
\end{definition}

In a construction by a Euclidean measure, these distributions are moments of
random distributions.  In an axiomatic Euclidean formulation, the hierarchy
\(\{S_n\}_{n\ge0}\) is primary data.

Let \(\theta\) be time reflection:
\[
  \theta(\tau,\mathbf x)=(-\tau,\mathbf x).
\]
The Euclidean group is
\[
  \operatorname{Euc}(D)=\mathbb R^D\rtimes O(D).
\]
It acts on test insertions by pullback on \(E^n\) and by the declared label
representation on \(V\).
The reflection also acts on labels.  We denote by \(J:V\to V\) the antilinear
label map induced by field adjunction and by the Euclidean reflection of
tensor or spinor indices.  For a scalar Hermitian field, \(J\) is ordinary
complex conjugation.  For a test function \(f_n\) supported in the ordered
positive-time region
\[
  E^n_>=\{(x_1,\ldots,x_n):0<\tau_1<\cdots<\tau_n\},
\]
define the reflected adjoint test function \(\Theta f_n\) by reversing the
order, reflecting all arguments, and applying \(J\) to the labels.  For a
scalar test function this reads
\[
  (\Theta f_n)(x_1,\ldots,x_n)
  =
  \overline{f_n(\theta x_n,\ldots,\theta x_1)}.
\]
With labels, the same formula is composed with \(J^{\otimes n}\).  Let
\(\Theta\) denote the same antilinear operation on arbitrary \(n\)-point test
functions; positive-time support is needed only when we form the OS
pre-Hilbert space below.  Let
\(\mathcal E_+\) be the vector space of finite sequences
\[
  F=(f_0,f_1,f_2,\ldots),
\]
where only finitely many \(f_n\) are nonzero and every \(f_n\) is a labelled
zero-diagonal test function supported in positive Euclidean time with a chosen
ordering of its time variables.  Reflection positivity is the requirement that the
sesquilinear form
\begin{equation}
  \langle F,G\rangle_{\rm OS}
  =
  \sum_{m,n} S_{m+n}\bigl(\Theta f_m,g_n\bigr)
  \label{eq:os-positive-time-form}
\end{equation}
satisfy
\begin{equation}
  \langle F,F\rangle_{\rm OS}\ge0
  \label{eq:os-reflection-positivity}
\end{equation}
for every \(F\in\mathcal E_+\).

The notation \(S_{m+n}(\Theta f_m,g_n)\) means that the reflected \(m\)
negative-time insertions are followed by the \(n\) positive-time insertions in
the argument of the distribution.  Thus the object required to be positive is
the reflected positive-time sesquilinear form on \(\mathcal E_+\).

\begin{definition}[OS-admissible Euclidean data]
\label{def:os-admissible-euclidean-data}
An OS-admissible Euclidean field presentation is a Schwinger hierarchy
\(\{S_n\}_{n\ge0}\), together with the label reflection \(J\), satisfying:
\begin{enumerate}[label=(\roman*)]
  \item normalization: \(S_0=1\);
  \item Euclidean covariance: \(S_n(gf_n)=S_n(f_n)\) for
  \(g\in \operatorname{Euc}(D)\);
  \item permutation symmetry: \(S_n\) is invariant under simultaneous
  permutation of arguments and labels;
  \item Hermiticity: for every \(n\ge0\) and every labelled \(n\)-point test
  function \(f_n\),
  \begin{equation}
    S_n(\Theta f_n)
    =
    \overline{S_n(f_n)} .
    \label{eq:os-hermiticity}
  \end{equation}
  \item reflection positivity:
  \(\langle F,F\rangle_{\rm OS}\ge0\) for all \(F\in\mathcal E_+\);
  \item positive-time semigroup regularity: the positive-time translation
  operators defined on the OS quotient form a strongly continuous symmetric
  contraction semigroup, as stated in
  \eqref{eq:os-time-semigroup-law}--\eqref{eq:os-time-strong-continuity};
  \item corrected OS-II analytic regularity: the difference-variable hierarchy
  satisfies a many-variable continuation and growth theorem strong enough to
  give the Lorentzian boundary-value package of
  Definition~\ref{def:os-lorentzian-boundary-value-package}.  A standard
  sufficient hypothesis is the OS-II linear-growth condition
  \eqref{eq:os-linear-growth-condition}, together with the OS-II
  several-complex-variable construction described below.
\end{enumerate}
For fermionic or graded fields, the permutation line is replaced by graded
permutation symmetry with the Koszul sign determined by the declared parity of
each label.  Clustering is added when the target Lorentzian theory is required
to have a unique vacuum vector.
\end{definition}

The last line is a load-bearing analytic hypothesis, not a cosmetic
regularity clause.  Let \(S_n^{\rm red}\) denote the translation-reduced
Schwinger distribution in \(n-1\) difference variables, acting on the
corresponding zero-diagonal difference-test space.  With a fixed family of
Schwartz seminorms
\[
  q_{n,N}(F)
  =
  \sum_{|\alpha|,|\beta|\le N}
  \sup_{\Xi\in E^{n-1}}
  \left|\Xi^\beta\partial^\alpha F(\Xi)\right|,
  \qquad F\in\mathcal S_\circ(E^{n-1};V^{\otimes n}),
\]
one convenient OS-II linear-growth condition is the existence of constants
\[
  C_{\rm OS},N_0\in\mathbb N,\qquad
  A,B,\gamma>0,
\]
independent of \(n\), such that
\begin{equation}
  |S_n^{\rm red}(F)|
  \le
  A\,B^n\,(n!)^\gamma\,
  q_{n,N_0+C_{\rm OS}n}(F)
  \label{eq:os-linear-growth-condition}
\end{equation}
for all \(F\) in that test space.  Equivalent OS-II formulations package the
same content by using a fixed seminorm family \(|\cdot|_{s,n}\) whose
definition already contains the \(n\)-dependence, together with constants
\(\sigma_n\le A B^n(n!)^\gamma\).  The word ``linear'' refers to the allowed
growth of the distributional seminorm order with the number of insertions.  It
is not a statement that a fixed \(n\)-point function grows linearly in
spacetime.

Together with the positive-time semigroup hypotheses
\eqref{eq:os-time-semigroup-law}--\eqref{eq:os-time-strong-continuity}, this
condition has a specific OS-II role.  The positive-time Hilbert space and the
one-variable semigroup continuation are constructed before the linear-growth
estimate is used.  The linear-growth estimate controls the many-variable
analytic continuation and the polynomial boundary estimates needed to obtain
tempered Lorentzian Wightman distributions.  The correction from the first
Osterwalder--Schrader reconstruction paper to the second is precisely that
reflection positivity, Euclidean covariance, symmetry, clustering, and ordinary
temperedness do not by themselves prove that the separately obtained
one-variable Laplace continuations assemble into a joint many-variable
Fourier--Laplace transform with tempered boundary values.
Thus an application of OS reconstruction must either prove an OS-II growth
condition such as \eqref{eq:os-linear-growth-condition}, or replace it by a
different theorem that supplies the same Lorentzian boundary-value estimates.
\footnote{See K.~Osterwalder and R.~Schrader, ``Axioms for Euclidean Green's
Functions,'' Commun. Math. Phys. \textbf{31} (1973) 83--112, and ``Axioms for
Euclidean Green's Functions II,'' Commun. Math. Phys. \textbf{42} (1975)
281--305.  The second paper states the corrected reconstruction theorem with
the linear-growth input.  Modern checks of this analytic input are model
dependent.  For example, Gubinelli and Hofmanov\'a, ``A PDE Construction of
the Euclidean \(\Phi^4_3\) Quantum Field Theory,'' Commun. Math. Phys.
\textbf{384} (2021) 1--75, construct limit measures with reflection
positivity, translation invariance, non-Gaussianity, and stretched exponential
integrability estimates strong enough for the OS distributional regularity
axiom in their formulation; rotation invariance and clustering remain separate
inputs in that construction.}

\begin{remark}[CFT reconstruction is a different route]
For a Euclidean CFT, one should not silently replace the OS-II growth package
by conformal terminology.  Conformal kinematics gives additional structure:
radial ordering, the inversion pairing, positivity of the state-operator
inner product, and conformal covariance of operator insertions.  In that
setting a Lorentzian theory can be reconstructed by proving the convergence
and distributional boundary-value properties of conformal correlation
functions directly from the conformal Hilbert-space and OPE data.  This is a
different theorem from OS reconstruction.  Kravchuk, Qiao, and Rychkov prove
the relevant Lorentzian distribution and OPE-convergence estimates for CFT
correlators using conformal kinematics and radial variables;\footnote{See
P.~Kravchuk, J.~Qiao, and S.~Rychkov, ``Distributions in CFT II. Minkowski
Space,'' JHEP \textbf{08} (2021) 094, arXiv:2104.02090.} the CFT volume uses
that route only after the state-operator Hilbert space and conformal
representation theory have been stated as hypotheses.
\end{remark}

Figure~\ref{fig:os-reflection-positivity} fixes the geometry used in the OS
reflection form.  In this Euclidean reconstruction figure the reflection time
\(\tau\) is drawn horizontally, so that the fixed plane \(\tau=0\) is the
vertical dashed line.

\begin{figure}[H]
  \centering
  \resizebox{0.82\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      axis/.style={-{Latex[length=2mm]},line width=0.75pt},
      dot/.style={circle,draw,fill=blue!12,line width=0.8pt,minimum size=0.34cm},
      rdot/.style={circle,draw,fill=orange!14,line width=0.8pt,minimum size=0.34cm},
      arrow/.style={-{Latex[length=2mm]},line width=0.85pt}
    }
    \draw[axis] (-4.2,0)--(4.2,0) node[right] {\(\tau\)};
    \draw[axis] (0,-2.0)--(0,2.0) node[above] {space};
    \draw[line width=0.8pt,dashed] (0,-1.75)--(0,1.75);
    \node[dot] (a) at (1.15,0.95) {};
    \node[dot] (b) at (2.55,-0.55) {};
    \node[rdot] (ta) at (-1.15,0.95) {};
    \node[rdot] (tb) at (-2.55,-0.55) {};
    \draw[arrow,orange!80!black] (a)--(ta)
      node[midway,above] {\(\theta\)};
    \draw[arrow,orange!80!black] (b)--(tb)
      node[midway,below] {\(\theta\)};
    \node[blue!55!black] at (2.15,1.55) {positive-time tests};
    \node[orange!80!black] at (-2.25,1.55) {reflected tests};
  \end{tikzpicture}
  }
  \caption{Reflection positivity turns Euclidean reflection through
  \(\tau=0\) into Hilbert-space positivity after reconstruction.  The
  positive quadratic form \(\sum S(\Theta f,f)\ge0\) descends, after quotient
  by null vectors, to the reconstructed Lorentzian inner product.}
  \label{fig:os-reflection-positivity}
\end{figure}

\section{Hilbert Space From Positive Time}

Use the positive-time vector space \(\mathcal E_+\) and the form
\eqref{eq:os-positive-time-form}.  Reflection positivity says that this form
is nonnegative.  Let
\[
  \mathcal N=\{F\in\mathcal E_+\mid \langle F,F\rangle_{\rm OS}=0\}.
\]
The reconstructed Hilbert space is
\[
  \Hilb_{\rm OS}
  =
  \overline{\mathcal E_+/\mathcal N}.
\]
The vector represented by \(F=(1,0,0,\ldots)\) is the vacuum.
The quotient is taken before completion.  Since
\eqref{eq:os-positive-time-form} is positive semidefinite, the
Cauchy--Schwarz inequality holds on \(\mathcal E_+\), and
\(\mathcal N\) is exactly the subspace of zero-norm vectors.  Therefore
\(\mathcal E_+/\mathcal N\) is a pre-Hilbert space with inner product induced
by \(\langle\cdot,\cdot\rangle_{\rm OS}\).

Euclidean time translations by \(s\ge0\) preserve positive-time support and
therefore act on \(\mathcal E_+\).  Write \(a_s(\tau,\mathbf x)=(\tau+s,\mathbf
x)\), and let \(R_s\) be the induced action on test sequences, so that each
component is translated by \(a_s\).  The OS regularity hypotheses needed at
this point are the following seminorm statements on \(\mathcal E_+\):
\begin{align}
  R_0&=\mathbf 1,
  &
  R_{s+t}&=R_sR_t,
  \label{eq:os-time-semigroup-law}\\
  \|R_sF\|_{\rm OS}&\le \|F\|_{\rm OS},
  \label{eq:os-time-contractivity}\\
  \langle R_sF,G\rangle_{\rm OS}
  &=
  \langle F,R_sG\rangle_{\rm OS},
  \label{eq:os-time-symmetry}\\
  \lim_{s\downarrow0}\|R_sF-F\|_{\rm OS}&=0,
  \label{eq:os-time-strong-continuity}
\end{align}
for all \(F,G\in\mathcal E_+\), where
\(\|F\|_{\rm OS}^2=\langle F,F\rangle_{\rm OS}\).  The inequality
\eqref{eq:os-time-contractivity} is the contraction statement needed to obtain
a nonnegative Hamiltonian.  In the classical scalar OS proof it is derived from
reflection positivity, Euclidean covariance, the semigroup law, and a
polynomial bound on translated matrix elements by the iterative
Cauchy--Schwarz argument written below.  In the abstract formulation of
OS-admissible data we record the contraction and strong continuity explicitly,
so that every later use of the Hamiltonian has a named hypothesis.

\begin{remark}[Failure mode of the first OS reconstruction theorem]
\label{rem:os-i-lemma-eight-eight-gap}
The famous error in the first Osterwalder--Schrader reconstruction paper is
not the positive-time Hilbert-space quotient and not the one-variable
semigroup continuation.  Those parts are the reliable core of the construction.
The faulty step was the passage from separately controlled one-variable
Laplace continuations to a joint many-variable Fourier--Laplace transform.  A
function on \((0,\infty)^2\) may be a Laplace transform of a tempered
distribution in each variable separately without being a joint Laplace
transform; the model obstruction is \(F(x,y)=\exp(-xy)\).  The corrected OS-II
proof replaces that step by a several-complex-variable continuation ladder and
then uses a linear-growth estimate to prove the polynomial bounds that make the
boundary values tempered distributions.\footnote{OS II states explicitly that
Lemma 8.8 of OS I is wrong and gives the \(F(x,y)=\exp(-xy)\) obstruction.  See
K.~Osterwalder and R.~Schrader, ``Axioms for Euclidean Green's Functions II,''
Commun. Math. Phys. \textbf{42} (1975) 281--305, Introduction and Chapters
IV--VI.}
\end{remark}

\begin{proposition}[Contraction from the OS semigroup estimate]
\label{prop:os-semigroup-contraction-iteration}
Let \(R_t:\mathcal E_+\to\mathcal E_+\), \(t\ge0\), satisfy the semigroup law
\(R_{t+s}=R_tR_s\), the OS symmetry relation
\[
  \langle R_tF,G\rangle_{\rm OS}
  =
  \langle F,R_tG\rangle_{\rm OS},
  \qquad F,G\in\mathcal E_+,
\]
and, for each \(F\in\mathcal E_+\), a polynomial growth bound
\[
  |\langle F,R_uF\rangle_{\rm OS}|
  \le C_F(1+u)^{N_F},
  \qquad u\ge0.
\]
Then
\[
  \|R_tF\|_{\rm OS}\le \|F\|_{\rm OS},
  \qquad t\ge0.
\]
In particular \(R_t\mathcal N\subset\mathcal N\), so \(R_t\) descends to a
contraction on the OS quotient.
\end{proposition}

\begin{proof}
Assume first that \(\|F\|_{\rm OS}>0\).  Put
\[
  a_j=\langle F,R_{2^jt}F\rangle_{\rm OS},
  \qquad j=0,1,2,\ldots .
\]
For \(j\ge1\), the semigroup law and OS symmetry give
\[
  a_j
  =
  \langle R_{2^{j-1}t}F,R_{2^{j-1}t}F\rangle_{\rm OS}
  \ge0 .
\]
The Cauchy--Schwarz inequality in the positive semidefinite OS form gives,
for \(j\ge1\),
\[
  a_j^2
  =
  |\langle F,R_{2^jt}F\rangle_{\rm OS}|^2
  \le
  \|F\|_{\rm OS}^2
  \|R_{2^jt}F\|_{\rm OS}^2
  =
  a_0\,a_{j+1}.
\]
Induction therefore yields
\[
  a_1
  \le
  a_0^{1-2^{-N}}a_N^{2^{-N}},
  \qquad N\ge1 .
\]
The polynomial bound gives
\[
  a_N^{2^{-N}}
  \le
  C_F^{2^{-N}}(1+2^Nt)^{N_F2^{-N}}
  \longrightarrow 1 .
\]
Letting \(N\to\infty\) gives \(a_1\le a_0\).  Since
\[
  a_1=\langle F,R_{2t}F\rangle_{\rm OS}
      =\langle R_tF,R_tF\rangle_{\rm OS}
      =\|R_tF\|_{\rm OS}^2,
\]
the contraction estimate follows.  If \(\|F\|_{\rm OS}=0\), Cauchy--Schwarz
already gives \(\langle F,R_uF\rangle_{\rm OS}=0\) for all \(u\ge0\), and the
same identity gives \(\|R_tF\|_{\rm OS}=0\).  Thus null vectors are preserved,
and the quotient operator is a contraction.
\end{proof}

Contractivity implies \(R_s\mathcal N\subset\mathcal N\), so \(R_s\) descends to
a bounded operator \(T_E(s)\) on the quotient and then to \(\Hilb_{\rm OS}\).
Equations
\eqref{eq:os-time-semigroup-law}--\eqref{eq:os-time-strong-continuity} say
precisely that \(\{T_E(s)\}_{s\ge0}\) is a strongly continuous semigroup of
self-adjoint contractions.
Indeed, \eqref{eq:os-time-symmetry} gives, on the dense quotient,
\[
  \langle [F],T_E(s)[G]\rangle_{\rm OS}
  =
  \langle [F],[R_sG]\rangle_{\rm OS}
  =
  \langle [R_sF],[G]\rangle_{\rm OS}
  =
  \langle T_E(s)[F],[G]\rangle_{\rm OS}.
\]
The equality descends to the quotient because \(R_s\mathcal N\subset
\mathcal N\), and it extends to \(\Hilb_{\rm OS}\) because \(T_E(s)\) is
bounded by \eqref{eq:os-time-contractivity}.  Thus the self-adjointness of
the Euclidean time semigroup is not inferred from a formal reflection
manipulation outside positive time; it is precisely the OS-II symmetry
condition on the reflection-positive quotient.

Let \(A\) be its infinitesimal generator,
\[
  A\psi=\lim_{s\downarrow0}\frac{T_E(s)\psi-\psi}{s},
\]
on the vectors for which the limit exists in \(\Hilb_{\rm OS}\).  The
self-adjoint contraction-semigroup theorem gives \(A=A^*\) and
\(\sigma(A)\subset(-\infty,0]\).  Setting \(H=-A\), the spectral theorem gives
\[
  T_E(s)=\ee^{-sH},\qquad H=H^*,\qquad H\ge0 .
\]
Stone's theorem then applies to the strongly continuous unitary group
\(U_{\rm time}(t)=\ee^{-\ii tH}\): its named hypotheses are the unitary group
law, strong continuity in \(t\), and self-adjointness of the generator.  This
\(H\) is the Lorentzian Hamiltonian in the reconstructed vacuum sector.

\begin{proposition}[One-gap continuation from the OS semigroup]
\label{prop:os-one-gap-semigroup-continuation}
Let \(F,G\in\mathcal E_+\) be two positive-time test sequences whose supports
are arranged so that, for \(s>0\), \(R_sG\) remains later than the insertions in
\(F\).  Let
\[
  \psi_F=[F],\qquad \psi_G=[G]\in\Hilb_{\rm OS}.
\]
Then the Euclidean matrix element with distinguished time gap \(s\) is
\[
  \langle \psi_F,T_E(s)\psi_G\rangle_{\rm OS}
  =
  \langle \psi_F,\ee^{-sH}\psi_G\rangle_{\Hilb_{\rm OS}} .
\]
For \(z\in\mathbb C\) with \(\Re z>0\), define
\[
  M_{F,G}(z)=\langle\psi_F,\ee^{-zH}\psi_G\rangle_{\Hilb_{\rm OS}} .
\]
Then \(M_{F,G}\) is holomorphic on the right half-plane and satisfies
\[
  |M_{F,G}(z)|\le \|\psi_F\|\,\|\psi_G\|,
  \qquad \Re z>0.
\]
Moreover, for every \(t\in\mathbb R\),
\[
  \lim_{\epsilon\downarrow0}M_{F,G}(\epsilon+\ii t)
  =
  \langle\psi_F,\ee^{-\ii tH}\psi_G\rangle_{\Hilb_{\rm OS}} .
\]
\end{proposition}

\begin{proof}
The first identity is the definition of the descended positive-time
translation semigroup:
\[
  T_E(s)[G]=[R_sG],
  \qquad T_E(s)=\ee^{-sH}.
\]
For the holomorphic extension, let \(E_H\) be the spectral resolution of the
nonnegative self-adjoint operator \(H\), and let
\[
  \mu_{F,G}(\Delta)=
  \langle \psi_F,E_H(\Delta)\psi_G\rangle
\]
be the finite complex spectral measure associated to \(\psi_F,\psi_G\).
Its total variation is bounded by
\(|\mu_{F,G}|([0,\infty))\le\|\psi_F\|\,\|\psi_G\|\).  The spectral theorem gives
\[
  M_{F,G}(z)=\int_{[0,\infty)}\ee^{-z\lambda}\,\dd\mu_{F,G}(\lambda),
  \qquad \Re z>0.
\]
On compact subsets of the right half-plane, derivatives with respect to \(z\)
are obtained by differentiating under the integral:
\[
  \partial_z^k M_{F,G}(z)
  =
  \int_{[0,\infty)}(-\lambda)^k\ee^{-z\lambda}\,
  \dd\mu_{F,G}(\lambda),
\]
because \(\lambda^k\ee^{-(\Re z)\lambda}\) is bounded on
\([0,\infty)\) uniformly for \(\Re z\) bounded below by a positive constant.
This proves holomorphy.  The norm bound follows from
\(|\ee^{-z\lambda}|\le1\).  Finally,
\(\ee^{-(\epsilon+\ii t)\lambda}\to\ee^{-\ii t\lambda}\) pointwise and is
bounded by \(1\); dominated convergence for the finite measure
\(|\mu_{F,G}|\) gives the stated boundary value.
\end{proof}

This proposition is the sound one-variable part of the OS construction.  It
does not by itself produce an \(n\)-variable Wightman distribution.  The
next QFT-specific task is to show that all one-gap continuations coming
from different cuts of the same Schwinger function assemble into a joint
many-variable analytic function with the spectral support and boundary-value
estimates required below.  That assembly is the content of the corrected OS-II
analytic theorem.

Spatial translations and rotations act unitarily on the quotient.
Together with the analytic continuation of Euclidean rotations, the full OS
theorem gives a strongly continuous unitary representation of the connected
Poincare group with joint translation spectrum contained in the closed forward
cone.

The passage from Euclidean covariance to Lorentzian covariance occurs here:
the Lorentzian Hilbert space and Poincare representation are constructed from
the positive-time quotient and the analytic-continuation data.

\begin{figure}[H]
  \centering
  \resizebox{0.94\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      box/.style={draw,rounded corners=4pt,line width=0.8pt,fill=blue!4,
        minimum width=3.25cm,minimum height=0.90cm,align=center},
      arrow/.style={-{Latex[length=2mm]},line width=0.85pt}
    }
    \node[box] (sch) at (-5.25,0.45) {Schwinger hierarchy\\\(\{S_n\}\)};
    \node[box] (pos) at (-1.75,0.45) {positive-time\\test sequences};
    \node[box] (hilb) at (1.75,0.45) {OS quotient\\\(\Hilb_{\rm OS}\)};
    \node[box] (lor) at (5.25,0.45)
      {Wightman data\\\(\Hilb_{\rm OS},\vac_{\rm OS},U,\widehat\Phi_\alpha\)};
    \node[box,fill=green!7] (semigroup) at (0,-1.45)
      {Euclidean time semigroup\\\(\ee^{-sH}\)};
    \draw[arrow] (sch)--(pos);
    \draw[arrow] (pos)--(hilb);
    \draw[arrow] (hilb)--(lor);
    \draw[arrow] (pos)--(semigroup);
    \draw[arrow] (semigroup)--(hilb);
    \node[blue!60!black] at (-3.50,1.32) {reflection positivity};
    \node[blue!60!black] at (3.50,1.32) {analytic continuation};
  \end{tikzpicture}
  }
  \caption{OS reconstruction turns Euclidean correlation data into a
  Lorentzian Hilbert-space theory through a positive-time quotient.}
  \label{fig:os-reconstruction-flow}
\end{figure}

\section{Fields on the Reconstructed Domain}

Let \(\mathcal D_{\rm OS}\subset\Hilb_{\rm OS}\) be the image of
\(\mathcal E_+\) in the quotient.  This dense subspace is the natural common
domain for the reconstructed fields before closure.  If \(h\) is a Euclidean
test function valued in \(V\), with support at positive time and ordered
before the supports appearing in \(F\in\mathcal E_+\), define
\[
  \Phi_E(h)[F]
  =
  [h\otimes F],
\]
where \(h\otimes F\) denotes the sequence obtained by inserting one additional
Euclidean field smeared with \(h\).  This formula is only meaningful after
one checks that it descends through the null quotient.  The check is a
reflection-positive adjointness argument, not a formal manipulation with
point fields.

\begin{proposition}[Well-definedness and closability of ordered insertions]
\label{prop:os-ordered-field-closability}
Assume the operator-domain part of the OS data supplies the following
adjoint-insertion identity.  For every ordered positive-time test insertion
\(h\) there is an ordered test insertion \(h^\sharp\), with adjoint label and
with supports arranged by a common positive time translation, such that for
all \(F,G\in\mathcal E_+\)
\begin{equation}
  \langle [G],[h\otimes F]\rangle_{\rm OS}
  =
  \langle [h^\sharp\otimes G],[F]\rangle_{\rm OS}.
  \label{eq:os-ordered-insertion-adjoint-identity}
\end{equation}
Then \(F\in\mathcal N_{\rm OS}\) implies
\(h\otimes F\in\mathcal N_{\rm OS}\).  Hence
\(\Phi_E(h)[F]=[h\otimes F]\) is a well-defined operator on the quotient
domain \(\mathcal D_{\rm OS}\).  This operator is closable.
\end{proposition}

\begin{proof}
Let \(F\in\mathcal N_{\rm OS}\).  In a positive semidefinite Hermitian form,
Cauchy--Schwarz gives, for every \(K\in\mathcal E_+\),
\[
  \langle K,F\rangle_{\rm OS}=0 .
\]
Therefore, for every \(G\in\mathcal E_+\),
\[
  \langle [G],[h\otimes F]\rangle_{\rm OS}
  =
  \langle [h^\sharp\otimes G],[F]\rangle_{\rm OS}
  =
  0 .
\]
Taking \(G=h\otimes F\), with the common positive time translation already
included in the ordered-support hypothesis, gives
\(\|h\otimes F\|_{\rm OS}^2=0\).  Thus \(h\otimes F\in\mathcal N_{\rm OS}\),
and the quotient operator is well defined.

Let \(\psi_k\in\mathcal D_{\rm OS}\) satisfy
\(\psi_k\to0\) and \(\Phi_E(h)\psi_k\to\eta\) in \(\Hilb_{\rm OS}\).  To
prove closability it suffices to show \(\eta=0\).  For a dense test vector
\([G]\in\mathcal D_{\rm OS}\), the adjoint-insertion identity gives
\[
  \langle [G],\eta\rangle_{\rm OS}
  =
  \lim_k\langle [G],\Phi_E(h)\psi_k\rangle_{\rm OS}
  =
  \lim_k\langle [h^\sharp\otimes G],\psi_k\rangle_{\rm OS}
  =
  0,
\]
because \([h^\sharp\otimes G]\) is a fixed vector in
\(\mathcal D_{\rm OS}\) and \(\psi_k\to0\).  Since \(\mathcal D_{\rm OS}\) is
dense, \(\eta\) is orthogonal to all of \(\Hilb_{\rm OS}\), hence
\(\eta=0\).  This is the standard closability criterion, with the
QFT-specific content isolated in the matrix-element identity
\eqref{eq:os-ordered-insertion-adjoint-identity}.
\end{proof}

Lorentzian fields are obtained by continuing these matrix elements on the
reconstructed Hilbert space.  For ordered imaginary parts
\[
  \epsilon_1>\epsilon_2>\cdots>\epsilon_n>0,
\]
the Wightman distributions are boundary values of the Schwinger functions
continued to
\begin{equation}
  \tau_j=\epsilon_j+\ii t_j,
  \qquad j=1,\ldots,n.
  \label{eq:os-wightman-boundary-values}
\end{equation}
The reconstructed operator \(\widehat\Phi_\alpha(t,\mathbf f)\) is the
operator-valued distribution whose vacuum matrix elements are these boundary
values and whose time dependence is implemented by the unitary group
\[
  \ee^{-\ii tH}.
\]
Thus the OS construction produces the vacuum vector, the Hilbert space, the
positive Hamiltonian, and the field domains from the same Euclidean
correlation hierarchy.

As in the Wightman presentation, these reconstructed fields are generally
unbounded distributional coordinates.  A represented local observable algebra
\(\mathcal R(\mathcal O)\) is obtained by specifying a closure, affiliation,
or bounded functional-calculus construction, or by building a local net
directly and proving that the reconstructed fields are affiliated with it.
The OS theorem reconstructs a Lorentzian Wightman field
presentation; comparison with AQFT local observable content is a further
construction with its own hypotheses.

\section{Analytic Continuation Input}

The OS-II regularity assumptions are used through the following
boundary-value package.  It is stated separately because this is the part of
the reconstruction theorem where the first OS paper needed correction.

\begin{definition}[OS Lorentzian boundary-value package]
\label{def:os-lorentzian-boundary-value-package}
For each \(n\), the Schwinger distribution \(S_n\) is said to satisfy the OS
Lorentzian boundary-value package if the following data and estimates are
available.

\begin{enumerate}[label=(\roman*)]
\item For every Euclidean time ordering there is a holomorphic function
\[
  \mathcal W_n(z_1,\ldots,z_n)
\]
on the corresponding forward tube in the difference variables
\(\operatorname{Im}(z_j-z_{j+1})\in V_+^\circ\), where \(V_+^\circ\) is the
open Lorentzian forward cone, whose restriction to ordered imaginary times
agrees with \(S_n\) away from coincident points.
\item The holomorphic functions obey tempered tube bounds: for every compact
subcone \(K\subset (V_+^\circ)^{n-1}\) and every positive lower bound on the
distance of the imaginary parts from the boundary of the cone, the boundary
pairings with Schwartz tests are bounded by finitely many Schwartz seminorms
with polynomial dependence on the reciprocal distance to the cone boundary.
\item The permuted Euclidean orderings give holomorphic functions on the
corresponding permuted tubes that agree on their common Euclidean real
submanifolds.  For spacelike separated adjacent insertions, the two
permuted tubes have a connected edge-of-the-wedge continuation through the
corresponding Jost domain, with the Koszul sign included for graded fields.
\end{enumerate}
\end{definition}

\begin{proposition}[Tempered boundary values from polynomial tube bounds]
\label{prop:polynomial-tube-bounds-boundary-values}
Let \(C\subset\mathbb R^d\) be an open convex cone and let
\[
  T_C=\mathbb R^d+\ii C
\]
be the corresponding tube.  Let \(F\) be holomorphic on \(T_C\).  Suppose that
for every closed subcone \(K\Subset C\) there are integers \(M,N\) and a
constant \(A_K\) such that
\begin{equation}
  |\langle F(\cdot+\ii y),f\rangle|
  \le
  A_K\,\operatorname{dist}(y,\partial C)^{-M}
  \sum_{|\alpha|,|\beta|\le N}
    \sup_{x\in\mathbb R^d}|x^\beta\partial^\alpha f(x)|
  \label{eq:polynomial-tube-bound-abstract}
\end{equation}
for all \(y\in K\) and all \(f\in\mathcal S(\mathbb R^d)\), whenever the
left hand side is defined by integration over \(x\).  Then \(F\) has a unique
tempered boundary value
\[
  \operatorname{bv}F\in\mathcal S'(\mathbb R^d)
\]
in the sense that \(F(\cdot+\ii y)\to\operatorname{bv}F\) in
\(\mathcal S'(\mathbb R^d)\) as \(y\to0\) inside any proper subcone of \(C\).
If, in addition, \(F\) is the Fourier--Laplace transform of a distribution
carried by the dual closed cone
\[
  C^*=\{p\in\mathbb R^d\mid p\cdot y\ge0\text{ for all }y\in C\},
\]
then the Fourier transform of
\(\operatorname{bv}F\) is carried by \(C^*\).
\end{proposition}

\begin{proof}
This is a mathematical input from the theory of analytic functionals and
tempered distributions.  One proves it by applying Cauchy's estimates in
subtubes to obtain uniform finite-order distribution bounds, representing the
tested functions by Fourier transforms, and using the Vladimirov
boundary-value theorem for holomorphic functions of polynomial growth in tube
domains.  The carrier-cone statement is the Paley--Wiener--Schwartz cone
theorem.  These are several-complex-variable and distribution-theoretic
theorems; the QFT-specific work in OS reconstruction is to prove their
hypotheses for the holomorphic functions constructed from the Schwinger
hierarchy.
\end{proof}

\begin{proposition}[OS-II analytic route to the boundary-value package]
\label{prop:os-ii-analytic-route-to-boundary-package}
Fix \(k=n-1\) and write the ordered difference variables as
\[
  \xi_j=x_{j+1}-x_j,\qquad j=1,\ldots,k .
\]
Assume the following OS-II analytic data have been constructed from the
Schwinger hierarchy.
\begin{enumerate}[label=(\roman*)]
\item There are scalar analyticity domains \(C_k^{(N)}\) in the complexified
time-difference variables and mixed Hilbert-vector domains \(D_m^{(N)}\) for
the OS vectors \(\Psi_m\), with
\[
  \bigcup_{N\ge0} C_k^{(N)}
  =
  \{(\zeta_1,\ldots,\zeta_k):\Re\zeta_j>0\ \text{for all }j\}.
\]
\item On each \(C_k^{(N)}\) there is a holomorphic representative
\(S_k^{(N)}\) of the difference-variable Schwinger function, and the
representatives agree on overlaps.
\item The Hilbert-vector fields \(\Psi_m(x,\zeta)\) are holomorphic on
\(D_m^{(N)}\) and satisfy the OS scalar-product formula
\begin{equation}
  \bigl\langle
    \Psi_r(x,\zeta),\Psi_s(y,\eta)
  \bigr\rangle_{\Hilb_{\rm OS}}
  =
  S_{r+s-1}^{(N)}
  \bigl(-\overline\zeta,\ x+y,\ \eta\bigr)
  \label{eq:os-ii-scalar-product-pattern}
\end{equation}
whenever both sides are defined.  The block
\(-\overline\zeta\) is the reflected left block, \(x+y\) is the real bridge
time, and \(\eta\) is the right block.
\item The local step \(A_N\Rightarrow P_N\) is obtained by Hilbert-space
Taylor reconstruction: the squared norm of the vector Taylor remainder is the
scalar Taylor remainder of the function on the right hand side of
\eqref{eq:os-ii-scalar-product-pattern}.  The step \(P_N\Rightarrow A_{N+1}\)
is obtained from the scalar products of these vector fields and extension to
the envelope of holomorphy defining \(C_k^{(N+1)}\).
\item The continued scalar functions obey a uniform polynomial tube estimate:
for every closed subcone \(K\Subset(V_+^\circ)^k\) and every
\(\delta>0\), if the imaginary parts of the Lorentzian difference variables lie
in \(K\) at distance at least \(\delta\) from its boundary, then
\begin{equation}
  |\langle S_k(\cdot+\ii y),f\rangle|
  \le
  C_{k,K}\,\delta^{-N_k}\,
  \sum_{|\alpha|,|\beta|\le N_k}
  \sup_\xi |\xi^\beta\partial^\alpha f(\xi)|
  \label{eq:os-ii-polynomial-tube-estimate}
\end{equation}
for all Schwartz tests \(f\) in the reduced difference variables.
\item Euclidean covariance identifies the product right-half-plane
continuation with the standard forward-tube continuation in Lorentzian
difference variables, and the corresponding statements hold for every
permutation of the field insertions, with the Koszul sign for graded labels.
\end{enumerate}
Then the Schwinger hierarchy satisfies the OS Lorentzian boundary-value package
of Definition~\ref{def:os-lorentzian-boundary-value-package}.
\end{proposition}

\begin{proof}
Items (i)--(iii) provide a single holomorphic function in the ordered
difference variables: the functions \(S_k^{(N)}\) agree on overlaps, and the
union of their domains is the full product right half-plane in the ordered
time gaps.  Item (vi) transports this function by Euclidean covariance to the
usual forward tube in Lorentzian difference variables.  This gives part (i) of
Definition~\ref{def:os-lorentzian-boundary-value-package}.

The estimate \eqref{eq:os-ii-polynomial-tube-estimate} is exactly the
tempered tube-bound input in part (ii) of the boundary-value package.  The
right hand side is a finite Schwartz seminorm, and the dependence on the
distance to the cone boundary is polynomial.  Therefore the family of
functionals \(f\mapsto\langle S_k(\cdot+\ii y),f\rangle\) is bounded in a
fixed finite-order tempered-distribution topology on each closed subtube.
	Proposition~\ref{prop:polynomial-tube-bounds-boundary-values} then gives
	tempered boundary values as \(y\to0\) inside the cone.

For a permutation of insertions, item (vi) gives the corresponding permuted
tube.  Euclidean permutation symmetry says that the permuted holomorphic
functions have the same values on their common Euclidean ordered real
submanifolds, after inserting the Koszul sign for odd labels.  The identity
theorem and the edge-of-the-wedge theorem applied at Jost configurations
therefore give the connected continuation through the common real edge.  This
is part (iii) of Definition~\ref{def:os-lorentzian-boundary-value-package}.
\end{proof}

\begin{theorem}[Corrected OS-II analytic theorem]
\label{thm:corrected-os-ii-analytic-theorem}
Let the Schwinger hierarchy satisfy normalization, Euclidean covariance,
graded permutation symmetry, Hermiticity, reflection positivity, the
positive-time semigroup regularity stated above, and the linear-growth
condition \eqref{eq:os-linear-growth-condition}.  Work componentwise in a
fixed finite field-label basis; graded signs enter only when two labelled
insertions are permuted.  Then the OS-II analytic data listed in
Proposition~\ref{prop:os-ii-analytic-route-to-boundary-package} are obtained
from the Schwinger hierarchy.  Hence the hierarchy satisfies the OS
Lorentzian boundary-value package.
\end{theorem}

\begin{proof}
We write the proof for scalar components.  Since \(V\) is finite dimensional in
the present field presentation, labelled components are handled by the same
argument applied to matrix-valued distributions; the label reflection \(J\) and
the Koszul signs only change the component bookkeeping.

Let \(k=n-1\).  Translation invariance lets us write the ordered Euclidean
\(n\)-point function in reduced variables
\[
  \xi=(\xi_1,\ldots,\xi_k),\qquad
  \xi_j=x_{j+1}-x_j\in E .
\]
On a fixed ordered component of the complement of the partial diagonals, write
\[
  \xi_j=(t_j,\mathbf x_j),\qquad t_j>0 ,
\]
after a Euclidean transformation if necessary.  The complex variables used
below are the time gaps
\[
  \zeta=(\zeta_1,\ldots,\zeta_k),\qquad
  C_+^k=\{\zeta\in\mathbb C^k:\Re\zeta_j>0\}.
\]
Spatial difference variables are kept as real parameters until the final
Euclidean covariance step; this is only a notational separation, not a
restriction on covariance.

\emph{Step 1: local one-direction continuation.}
Choose a split of the insertions into a reflected left block and a positive
right block, and separate the two blocks by one positive Euclidean time gap
\(s>0\).  Reflection positivity gives OS vectors \(\psi_F,\psi_G\) and
Proposition~\ref{prop:os-one-gap-semigroup-continuation} gives
\[
  s\longmapsto
  S_{r+s-1}(-\xi_L,s,\xi_R)
  =
  \langle \psi_F,\ee^{-sH}\psi_G\rangle_{\Hilb_{\rm OS}}
\]
as the boundary value on \(s>0\) of a function holomorphic for
\(\Re z>0\).  Euclidean covariance lets us rotate the distinguished positive
direction.  Thus every noncoincident real configuration has finitely many
one-complex-variable continuations in directions spanning the complexified
difference-variable space.

The passage from these one-variable continuations to a local
many-variable continuation is an application of the Malgrange--Zerner
extension theorem for compatible holomorphic continuations from flat wedges.
Explicitly, choose \(D\) real directions \(e_\mu\) which form a Euclidean
basis and a radius \(\rho(\xi)>0\) small enough that every displacement
\[
  \xi_j+\sum_{\mu=1}^D z_{j\mu}e_\mu,
  \qquad |z_{j\mu}|<\rho(\xi),
\]
stays in the same noncoincident ordered component after the relevant
rotations.  The one-variable continuations agree on their common real slices
because they all restrict to the same Schwinger distribution.  The
Malgrange--Zerner theorem therefore gives a holomorphic function in the
polydisc \(|z_{j\mu}|<c\rho(\xi)\), whose real restriction is the original
Schwinger function.  This proves real analyticity away from the partial
diagonals and gives the initial assertion \(A_0\).

\emph{Step 2: the \(A_N/P_N\) induction.}
We now define the induction precisely.  \(A_N(k)\) means that the reduced
\(k\)-gap Schwinger function has a holomorphic representative
\[
  S_k^{(N)}(\zeta\mid\mathbf x)
\]
for \(\zeta\in C_k^{(N)}\), continuous in the real spatial differences
\(\mathbf x=(\mathbf x_1,\ldots,\mathbf x_k)\), and agreeing with the
previous representatives on overlaps.  The initial domain \(C_k^{(0)}\)
is the union of the local polydiscs supplied by Step 1 around positive real
time gaps.

Given \(C_\ell^{(N)}\), define the vector domain \(D_m^{(N)}\) as the set of
pairs \((x,\zeta)\), with \(x>0\) a real bridge time and
\(\zeta=(\zeta_1,\ldots,\zeta_{m-1})\), for which the reflected norm argument
\[
  (-\overline\zeta_{m-1},\ldots,-\overline\zeta_1,\;2x,\;
    \zeta_1,\ldots,\zeta_{m-1})
\]
lies in \(C_{2m-1}^{(N)}\).  \(P_N(m)\) means that there is a holomorphic
\(\Hilb_{\rm OS}\)-valued function \(\Psi_m(x,\zeta)\) on this domain such
that, whenever the displayed scalar argument lies in \(C_{r+s-1}^{(N)}\),
  \begin{equation}
  \bigl\langle
    \Psi_r(x,\zeta),\Psi_s(y,\eta)
  \bigr\rangle_{\Hilb_{\rm OS}}
  =
  S_{r+s-1}^{(N)}
  (-\overline\zeta,\ x+y,\ \eta).
	  \label{eq:os-induction-scalar-product}
  \end{equation}
This is the scalar-product formula
\eqref{eq:os-ii-scalar-product-pattern}.

The implication \(A_N\Rightarrow P_N\) is the Hilbert Taylor argument.  Fix a
real point \((x,\xi)\in D_m^{(N)}\) and a polydisc \(P\Subset D_m^{(N)}\).
At the real point the vector \(\Psi_m(x,\xi)\) is the OS class of the ordered
test insertion.  Its strong derivatives are defined by differentiating the
test insertion before passing to the OS quotient.  For
\(\zeta\in P\) set the \(L\)-th vector Taylor remainder
\[
  R_L(\zeta)
  =
  \Psi_m(x,\zeta)
  -
  \sum_{|\alpha|\le L}
    \frac{(\zeta-\xi)^\alpha}{\alpha!}\,
    \partial^\alpha_\xi\Psi_m(x,\xi).
\]
Using \eqref{eq:os-induction-scalar-product} for real \(\xi\) and then
\(A_N(2m-1)\), the squared norm of this remainder is the scalar Taylor
remainder of the holomorphic function
\[
  S_{2m-1}^{(N)}(-\overline\zeta,2x,\zeta)
\]
about \((-\xi,2x,\xi)\).  The scalar Taylor remainder tends to zero uniformly
on smaller polydiscs.  Hence \(R_L(\zeta)\to0\) in
\(\Hilb_{\rm OS}\), and the Taylor series defines a holomorphic vector
\(\Psi_m(x,\zeta)\).  Polarizing the norm identity gives
\eqref{eq:os-induction-scalar-product}, so \(P_N\) follows.

Conversely, assume \(P_N(r)\) and \(P_N(s)\) for all lower vector lengths.
For \(k=r+s-1\) the scalar product
\[
  (x,\zeta;y,\eta)\longmapsto
  \langle\Psi_r(x,\zeta),\Psi_s(y,\eta)\rangle_{\Hilb_{\rm OS}}
\]
is holomorphic on the mixed vector domain and agrees with \(S_k\) on the real
ordered slice.  Let \(C_k^{(N+1)}\) be the envelope of holomorphy of the union
of all scalar domains generated in this way, together with \(C_k^{(N)}\).
The identity theorem gives a single extension \(S_k^{(N+1)}\).  Thus
\(P_N\Rightarrow A_{N+1}\).  Starting from \(A_0\), induction gives
\[
  A_0\Rightarrow P_0\Rightarrow A_1\Rightarrow P_1\Rightarrow\cdots .
\]

\emph{Step 3: exhaustion of the product half-plane.}
The induction just defined enlarges only the time-gap domains.  To see that it
reaches all of \(C_+^k\), use logarithmic coordinates
\[
  \zeta_j=\ee^{u_j+\ii v_j},\qquad -\frac{\pi}{2}<v_j<\frac{\pi}{2}.
\]
Let \(c_k^{(N)}\) be the closed set of allowed argument vectors
\((v_1,\ldots,v_k)\) for \(C_k^{(N)}\), and let \(d_m^{(N)}\) be the
corresponding set for \(D_m^{(N)}\).  The envelope-of-holomorphy theorem for
tubes says that the passage to \(C_k^{(N+1)}\) replaces the union of the
argument bases by its convex hull.  In these variables the recurrence is
\begin{equation}
  c_k^{(N+1)}
  =
  \operatorname{conv}
  \{(-v,\vartheta,w):
    (0,v)\in d_r^{(N)},\ (0,w)\in d_s^{(N)},\
    r+s-1=k,\ |\vartheta|\le\frac{\pi}{2}\},
  \label{eq:os-ii-argument-convex-recurrence}
\end{equation}
with the evident deletion of empty blocks.  The vector domains satisfy the
same recurrence with one distinguished real bridge argument fixed to \(0\).

The recurrence \eqref{eq:os-ii-argument-convex-recurrence} implies the
following elementary convex-geometry fact.  For every \(k\) and every
\(\epsilon>0\) there is an \(N(k,\epsilon)\) such that
\begin{equation}
  [-\frac{\pi}{2}+\epsilon,\frac{\pi}{2}-\epsilon]^k
  \subset c_k^{(N(k,\epsilon))}.
  \label{eq:os-ii-domain-exhaustion-box}
\end{equation}
Indeed, \(c_1^{(1)}\) contains the full interval
\([-\pi/2,\pi/2]\) because the single gap can be chosen as the bridge
variable.  If boxes of width \(a<\pi/2\) are available for all lower block
lengths, the recurrence forms convex hulls of points obtained by placing an
arbitrary bridge angle \(\vartheta\in[-\pi/2,\pi/2]\) between two already
available blocks and by reflecting the left block.  Averaging the \(k\)
possible placements of the bridge angle expands the available symmetric box by
a fixed positive fraction of the remaining distance to \(\pi/2\).  Iteration
therefore gives an increasing sequence \(a_N\uparrow\pi/2\), which proves
\eqref{eq:os-ii-domain-exhaustion-box}.  Since every point of \(C_+^k\) has
arguments strictly between \(-\pi/2\) and \(\pi/2\), this proves
\[
  \bigcup_{N\ge0}C_k^{(N)}=C_+^k .
\]

\emph{Step 4: real-domain polynomial bounds from linear growth.}
The preceding steps prove holomorphic continuation but not tempered boundary
values.  The linear-growth condition enters here.  Fix a real noncoincident
configuration \(\xi\) and let \(\rho(\xi)\) be the local analyticity radius
from Step 1, chosen so that
\begin{equation}
  \rho(\xi)\ge c_k(1+\|\xi\|)^{-a_k}
  \operatorname{dist}(\xi,\Delta)^{b_k}
  \label{eq:os-local-radius-lower-bound}
\end{equation}
on each ordered component; the exponents depend only on \(k\) and the
dimension.  Choose a nonnegative \(C^\infty\) radial kernel \(g_\rho\) on
\(\mathbb C^D\) with support in \(|z|<\rho/8\), total integral \(1\), and
derivative bounds
\[
  \|\partial^\alpha g_\rho\|_\infty
  \le C_\alpha \rho^{-2D-|\alpha|}.
\]
Let \(k_\rho=g_\rho*g_\rho\), so \(k_\rho\) is supported in
\(|z|<\rho/4\) and has total integral \(1\).  Applying the mean-value theorem
for holomorphic functions blockwise and inserting \(k_\rho\), one obtains a
regularized function \(T_{k,\rho}\) such that for \(|\eta_j|<\rho/2\)
\begin{equation}
  S_k(\xi+\eta)
  =
  \int T_{k,\rho}(\xi+\eta+\ii y)\,\dd\nu_\rho(y),
  \label{eq:os-regularized-mean-value}
\end{equation}
where \(\nu_\rho\) is a positive probability measure supported in
\(|y_j|<\rho/4\).  Consequently
\begin{equation}
  |S_k(\xi+\eta)|
  \le
  \sup_{|y_j|<\rho/4}|T_{k,\rho}(\xi+\eta+\ii y)|.
  \label{eq:os-regularized-sup-bound}
\end{equation}

The regularized function is designed so that OS positivity still applies.  For
each split \(k=r+s-1\) and each of the finitely many Euclidean basis
directions, the value of \(T_{k,\rho}\) after a real displacement in that
direction can be written as
\begin{equation}
  T_{k,\rho}(\xi+u e_\mu)
  =
  \langle \Phi_{L,\rho}(\xi,u),\Phi_{R,\rho}(\xi,u)\rangle_{\Hilb_{\rm OS}},
  \label{eq:os-regularized-scalar-product}
\end{equation}
where the two OS vectors are obtained by smearing the left and right ordered
blocks with the kernels \(g_\rho\) and \(k_\rho\).  This is an identity of
test functions before quotienting: the reflected left test and the right test
have exactly the convolution kernel appearing in \(T_{k,\rho}\).  The
semigroup continuation in the distinguished direction gives
\begin{equation}
  |T_{k,\rho}(\xi+w e_\mu)|
  \le
  \|\Phi_{L,\rho}(\xi,u)\|_{\Hilb_{\rm OS}}\,
  \|\Phi_{R,\rho}(\xi,u)\|_{\Hilb_{\rm OS}},
  \qquad \Re w>0 .
  \label{eq:os-regularized-semigroup-bound}
\end{equation}

The norm of, say, \(\Phi_{L,\rho}\) is an OS quadratic form
\(S(\Theta F,F)\) with an explicitly known test function \(F\).  Apply
\eqref{eq:os-linear-growth-condition} to this test.  The support bounds of
\(g_\rho\) and \(k_\rho\) imply that every spacetime argument in the quadratic
form is bounded by a polynomial in \(1+\|\xi\|+\|u\|\), while the derivative
bounds of the kernels contribute only powers of \(\rho^{-1}\).  Therefore
there are constants \(A_k,B_k,M_k\), with at most factorial growth in \(k\),
such that
\begin{equation}
  \|\Phi_{L,\rho}(\xi,u)\|_{\Hilb_{\rm OS}}\,
  \|\Phi_{R,\rho}(\xi,u)\|_{\Hilb_{\rm OS}}
  \le
  A_kB_k\,\rho^{-M_k}(1+\|\xi\|+\|u\|)^{M_k}.
  \label{eq:os-regularized-vector-norm-bound}
\end{equation}
Combining
\eqref{eq:os-regularized-semigroup-bound} and
\eqref{eq:os-regularized-vector-norm-bound} gives polynomial bounds in each
direction.  Passing to logarithmic variables turns the directional wedge
domains into flat tubes.  The Malgrange--Zerner theorem glues the directional
continuations, and the maximum principle on the envelope of holomorphy
propagates the same bound to the whole local polydisc.  Returning through
\eqref{eq:os-regularized-sup-bound} gives a real-domain estimate of the form
\begin{equation}
  |S_k(\xi)|
  \le
  A'_k\,\rho(\xi)^{-M'_k}(1+\|\xi\|)^{M'_k}.
  \label{eq:os-real-domain-tempered-estimate}
\end{equation}
Together with \eqref{eq:os-local-radius-lower-bound}, this is a polynomial
growth estimate on every ordered real component, with finite distributional
order controlled linearly in \(k\) by
\eqref{eq:os-linear-growth-condition}.

\emph{Step 5: continuation of the estimates to \(C_+^k\).}
It remains to carry the real-domain estimate through the
\(A_N/P_N\) ladder.  Let \(\mathcal B_{k,t}\) be the Banach space of continuous
functions of the \(k\) spatial difference variables with norm
\[
  \|f\|_{k,t}
  =
  \sup_{\mathbf x}
  (1+\max_j|\mathbf x_j|)^{-t}|f(\mathbf x)|.
\]
For \(t\) larger than the exponent in
\eqref{eq:os-real-domain-tempered-estimate}, the functions
\(\zeta\mapsto S_k^{(N)}(\zeta\mid\cdot)\) are
\(\mathcal B_{k,t}\)-valued holomorphic functions on \(C_k^{(N)}\).  Introduce
a regulator \(\epsilon>0\) and divide by a positive scalar majorant
\(Q_{k,\epsilon}(\zeta)\) whose logarithm is convex in the variables
\(\Re\zeta_j\):
\[
  S_{k,\epsilon}^{(N)}(\zeta)
  =
  Q_{k,\epsilon}(\zeta)^{-1}
  S_k^{(N)}(\zeta+\epsilon{\bf 1}\mid\cdot).
\]
The majorant is chosen so that the real-domain estimate gives
\begin{equation}
  \|S_{k,\epsilon}^{(0)}(\zeta)\|_{k,t}\le A_k
  \qquad(\zeta\in C_k^{(0)}),
  \label{eq:os-renormalized-real-bound}
\end{equation}
uniformly in \(\epsilon\).

Assume the bound is known on \(C_\ell^{(N)}\) for all relevant lower scalar
lengths.  On a mixed domain generated by \(P_N\), the scalar-product formula
and Cauchy--Schwarz give
\begin{equation}
  \|S_{r+s-1}^{(N)}(-\overline\zeta,x+y,\eta)\|_{r+s-1,t}
  \le
  \|S_{2r-1}^{(N)}(-\overline\zeta,2x,\zeta)\|_{2r-1,t}^{1/2}
  \|S_{2s-1}^{(N)}(-\overline\eta,2y,\eta)\|_{2s-1,t}^{1/2}.
  \label{eq:os-banach-cauchy-schwarz}
\end{equation}
The convexity of \(\log Q_{k,\epsilon}\) is precisely the algebraic statement
that the denominator for the left side is dominated by the product of the two
denominators on the right.  Therefore
\[
  \|S_{k,\epsilon}^{(N)}\|_{k,t}\le A_k
\]
on the generating mixed region.  Since \(C_k^{(N+1)}\) is the envelope of
holomorphy of that region and the functions take values in the Banach space
\(\mathcal B_{k,t}\), the Banach-valued maximum principle extends the same
bound to all of \(C_k^{(N+1)}\).  Induction gives the estimate on every
\(C_k^{(N)}\).

Finally choose \(N=N(\zeta)\) using the exhaustion
\eqref{eq:os-ii-domain-exhaustion-box}; the required \(N\) is bounded by a
constant times a logarithm of the inverse angular distance of \(\zeta\) to
\(\partial C_+^k\).  Undoing the harmless regulator \(\epsilon\) and the
majorant \(Q_{k,\epsilon}\) yields the tube estimate
\[
  |\langle S_k(\cdot+\ii y),f\rangle|
  \le
  C_{k,K}\,\delta^{-N_k}
  \sum_{|\alpha|,|\beta|\le N_k}
    \sup_\xi |\xi^\beta\partial^\alpha f(\xi)|
\]
for \(y\) in any closed subcone \(K\Subset(V_+^\circ)^k\) at distance
\(\delta\) from the cone boundary.  This is
\eqref{eq:os-ii-polynomial-tube-estimate}.  Euclidean covariance transports
the product right-half-plane continuation to the Lorentzian forward tubes, and
permutation symmetry supplies the permuted tubes with the graded signs.
Thus all hypotheses of
Proposition~\ref{prop:os-ii-analytic-route-to-boundary-package} have been
constructed from the stated Euclidean data and the linear-growth condition.
\end{proof}

\begin{proposition}[Fourier--Laplace boundary values from spectral support]
\label{prop:fourier-laplace-tempered-boundary-values}
Let \(d=D(n-1)\), let \(\Gamma\subset\mathbb R^d\) be a closed convex cone,
and let \(\Gamma^\circledast\) be its open dual cone:
\[
  \Gamma^\circledast
  =
  \{y\in\mathbb R^d\mid p\cdot y>0
    \text{ for every }0\ne p\in\Gamma\}.
\]
Let \(W\in\mathcal S'(\mathbb R^d)\) have Fourier transform
\(\widehat W\) supported in \(\Gamma\).  For \(y\in\Gamma^\circledast\)
define, by using any smooth cutoff equal to \(1\) on a conic neighbourhood of
\(\Gamma\) on which \(p\cdot y\) is nonnegative,
\begin{equation}
  \mathcal W(x+\ii y)
  =
  (2\pi)^{-d}
  \bigl\langle \widehat W(p),
    \exp(-\ii p\cdot x)\exp(-p\cdot y)\bigr\rangle .
  \label{eq:fourier-laplace-tube-transform}
\end{equation}
Then \(\mathcal W\) is holomorphic on the tube
\(\mathbb R^d+\ii\Gamma^\circledast\).  If \(K\Subset\Gamma^\circledast\)
is a closed subcone and \(\operatorname{dist}(y,\partial\Gamma^\circledast)
\ge\delta>0\), then for every Schwartz test function \(f\)
\[
  \left|
    \int_{\mathbb R^d}\mathcal W(x+\ii y) f(x)\,\dd^dx
  \right|
  \le
  C_{K,N}\,\delta^{-N}\,
  \sum_{|\alpha|,|\beta|\le N}
    \sup_p |p^\beta\partial_p^\alpha\widehat f(p)|
\]
for some \(N\) determined by the distributional order of \(\widehat W\).
Moreover,
\[
  \mathcal W(\cdot+\ii y)\longrightarrow W
  \qquad\text{in }\mathcal S'(\mathbb R^d)
\]
as \(y\to0\) inside \(K\).
\end{proposition}

\begin{proof}
Because \(\widehat W\) is a tempered distribution, there are \(C,N\) such
that
\[
  |\langle\widehat W,\psi\rangle|
  \le
  C\sum_{|\alpha|,|\beta|\le N}
  \sup_p |p^\beta\partial_p^\alpha\psi(p)|
  \qquad \psi\in\mathcal S(\mathbb R^d).
\]
For \(y\in\Gamma^\circledast\), the multiplier
\(\exp(-p\cdot y)\) is exponentially decreasing on the support cone
\(\Gamma\).  To make the pairing literally a Schwartz pairing, choose a
smooth conic cutoff \(\chi_y\) equal to \(1\) near \(\Gamma\) and supported
where \(p\cdot y\ge c_y|p|-C_y\); then
\(\chi_y\exp(-p\cdot y)\exp(-\ii p\cdot x)\) is a Schwartz multiplier after
testing in \(x\).  Different choices of \(\chi_y\) give the same value,
because their difference vanishes in a neighbourhood of
\(\operatorname{supp}\widehat W\).  Derivatives in \(x+\ii y\) only multiply
the cutoff multiplier by polynomials in \(p\), so the same estimate proves
holomorphy by differentiation under the distributional pairing.

If \(y\) stays in a closed subcone at distance \(\delta\) from the boundary,
one cutoff may be chosen uniformly on that subcone, and cone duality gives
\[
  p\cdot y\ge c_K\delta |p|
  \qquad (p\in\Gamma)
\]
with \(c_K>0\).  Hence every derivative of
\(\exp(-p\cdot y)\widehat f(p)\) is bounded by a finite linear combination
of Schwartz seminorms of \(\widehat f\) times a power of \(\delta^{-1}\).
This gives the displayed tube bound.

For the boundary value, test against \(f\).  The pairing equals
\[
  (2\pi)^{-d}
  \langle\widehat W(p), \ee^{-p\cdot y}\widehat f(p)\rangle .
\]
More explicitly, the left side is computed with the same cutoff convention,
and this convention is invisible to \(\widehat W\).  The cutoff functions
\(\chi\,\ee^{-p\cdot y}\widehat f(p)\) converge to \(\chi\,\widehat f(p)\)
in the Schwartz topology on every conic neighbourhood of \(\Gamma\) on which
\(\chi=1\) near \(\Gamma\): the exponential is bounded by \(1\) on
\(\Gamma\), the derivative terms are controlled by the rapid decrease of
\(\widehat f\), and the factors \(\ee^{-p\cdot y}-1\) converge uniformly on
compact \(p\)-sets while the tails are uniformly small.  Continuity of
\(\widehat W\) gives convergence to
\((2\pi)^{-d}\langle\widehat W,\widehat f\rangle=W(f)\).
\end{proof}

The linear-growth estimate \eqref{eq:os-linear-growth-condition} is one
standard route to this package in the OS-II theorem: the controlled growth of
distributional seminorm order is what keeps the Fourier--Laplace continuation
inside the tempered distribution class as the imaginary parts approach the
Lorentzian boundary.  In a concrete constructive model, this route is a
separate analytic estimate to be proved, not a formal consequence of
reflection positivity alone.

Applied to reconstruction, Proposition~\ref{prop:fourier-laplace-tempered-boundary-values}
is used with \(d=D(n-1)\), \(\Gamma=(\overline V_+)^{n-1}\), and \(W\) equal
to the translation-reduced Wightman distribution.  The OS-II linear-growth
condition is the Euclidean-side estimate that makes the reconstructed
matrix elements have the required finite distributional order uniformly in
the number of insertions.  The spectral support itself comes from the
nonnegative Hamiltonian, spatial translation covariance, and the
OS-continued Lorentz transformations.  Thus the corrected OS theorem has two
distinct analytic duties: it must construct the positive-energy
translation representation, and it must prove the polynomial tube bounds
needed for the boundary-value theorem above.

\begin{proposition}[Consequences of the OS boundary-value package]
\label{prop:os-boundary-package-consequences}
Assume Definition~\ref{def:os-lorentzian-boundary-value-package}.  Then:
\begin{enumerate}[label=(\alph*)]
\item The limits of \(\mathcal W_n(x+\ii y)\) as
\(y\to0\) within the ordered tube exist in
\(\mathcal S'((\mathbb M^D)^n)\) and define tempered Wightman distributions.
\item Permutation symmetry of the Schwinger hierarchy gives graded
commutativity of the boundary-value fields at spacelike separation.
\item Continuing the Wightman distributions back to ordered imaginary time
recovers the original \(S_n\) away from the partial diagonals.
\end{enumerate}
\end{proposition}

\begin{proof}
For (a), fix one ordering and write the reduced difference variables as
\(\xi=(\xi_1,\ldots,\xi_{n-1})\).  The tube bound in
Definition~\ref{def:os-lorentzian-boundary-value-package} says that, for
each smaller cone of imaginary parts, the function
\(\xi\mapsto\mathcal W_n(\xi+\ii y)\) defines a continuous functional on
\(\mathcal S(\mathbb R^{D(n-1)})\) with seminorm constants polynomial in the
distance of \(y\) to the cone boundary.  Proposition~\ref{prop:fourier-laplace-tempered-boundary-values}
displays the model estimate: after Fourier transformation, the imaginary
part inserts exponential damping against a distribution whose spectral
support lies in the forward cone.  Hence, for every Schwartz test \(f\), the
pairings
\(\langle\mathcal W_n(\cdot+\ii y),f\rangle\) have a limit as
\(y\to0\) within the ordered cone.  The limiting functional is continuous in
the same finite collection of Schwartz seminorms and is therefore a tempered
distribution.  This distribution is the ordered Wightman distribution.

For (b), first treat unsmeared configurations away from diagonals.  In
reduced variables, a real point \(\xi\) is a Jost point if every nonzero
positive linear combination of the successive differences
\(\xi_j=x_j-x_{j+1}\) is spacelike.  Such points are precisely the real
edges at which the forward tube for one ordering and the forward tube for a
permuted ordering meet after complex Lorentz transformations.  Euclidean
permutation symmetry gives equality, with the Koszul sign, of the two
holomorphic functions on the Euclidean real slice where both orderings are
defined.  By the identity theorem this equality holds in the common analytic
continuation region.  The edge-of-the-wedge theorem applies because the two
tempered boundary values have a common real edge at the Jost set and agree
there as distributions.  Therefore their analytic continuations coincide in
the envelope of holomorphy of the union of the two tubes.

Now smear the adjacent fields with Lorentzian test functions whose supports
are spacelike separated.  A partition of unity reduces the support of the
smearing to coordinate neighbourhoods whose relative configurations lie in
Jost neighbourhoods.  The boundary-value equality obtained above can then be
paired with these test functions.  The result is
\[
  \langle\vac,
    \cdots \widehat\Phi_\alpha(f)\widehat\Phi_\beta(g)\cdots\vac\rangle
  =
  (-1)^{|\alpha||\beta|}
  \langle\vac,
    \cdots \widehat\Phi_\beta(g)\widehat\Phi_\alpha(f)\cdots\vac\rangle
\]
for every surrounding finite string of test insertions.  Since the
finite-insertion vectors are dense in \(\mathcal D_{\rm OS}\), this is
exactly local graded commutativity on the reconstructed common domain.

For (c), let \(0<\tau_1<\cdots<\tau_n\) be an ordered Euclidean
configuration away from \(\Delta_n\).  The original Schwinger function and
the function obtained by first taking the Lorentzian boundary value and then
substituting \(t_j=-\ii\tau_j\) are restrictions of the same holomorphic
function \(\mathcal W_n\) on the connected ordered tube.  If two such
holomorphic functions had the same Lorentzian boundary value but differed in
the ordered Euclidean interior, their difference would have zero
distributional boundary value on a nonempty real edge and hence, by the same
edge-of-the-wedge uniqueness theorem, would vanish in the connected tube.
Thus the continuation back to ordered imaginary time recovers the original
\(S_n\) away from the partial diagonals.  Contact-term extensions on the
diagonals are additional Euclidean data and are not recovered from
off-diagonal analytic continuation alone.
\end{proof}

\section{Reconstruction Theorem}

\begin{theorem}[Osterwalder--Schrader reconstruction]
\label{thm:os-reconstruction}
Let \(\{S_n\}_{n\ge0}\) be Euclidean distributions satisfying Euclidean
covariance, permutation symmetry, reflection positivity, Hermiticity,
normalization, the OS-II positive-time semigroup conditions
\eqref{eq:os-time-semigroup-law}--\eqref{eq:os-time-strong-continuity}, and
the linear-growth condition \eqref{eq:os-linear-growth-condition}.  Assume
also the ordered-insertion adjoint identity
\eqref{eq:os-ordered-insertion-adjoint-identity} for the field labels to be
reconstructed.  By the
corrected OS-II analytic theorem,
Theorem~\ref{thm:corrected-os-ii-analytic-theorem}, these hypotheses supply
the analytic route of
Proposition~\ref{prop:os-ii-analytic-route-to-boundary-package}.  Then there exists
Wightman field data
\[
  (\Hilb_{\rm OS},\vac_{\rm OS},U,\mathcal D_{\rm OS},
  \{\widehat\Phi_\alpha\}_{\alpha\in I})
\]
whose Schwinger functions are \(\{S_n\}\) at noncoincident Euclidean points.
The reconstructed Wightman functions are obtained as boundary values of
analytic continuations of the \(S_n\).
The symbol \(\vac_{\rm OS}\) denotes the vacuum vector produced by this
reconstruction; it is a subscripted instance of the global vacuum notation
\(\vac\).
\end{theorem}

\begin{proof}
Step 1 constructs the Hilbert space.  Let \(\mathcal E_+\) be the vector
space of finite labelled Euclidean insertions whose time coordinates are
strictly positive.  Reflection positivity makes
\[
  \langle F,G\rangle_{\rm OS}:=S_{m+n}(\Theta F,G)
\]
a positive semidefinite Hermitian form on \(\mathcal E_+\).  Its null space is
\[
  \mathcal N_{\rm OS}
  =
  \{F\in\mathcal E_+:\langle F,F\rangle_{\rm OS}=0\}.
\]
The Cauchy--Schwarz inequality for positive semidefinite forms implies that
\(\langle F,G\rangle_{\rm OS}=0\) whenever \(F\in\mathcal N_{\rm OS}\);
therefore the form descends to \(\mathcal E_+/\mathcal N_{\rm OS}\).
Completing gives
\(\Hilb_{\rm OS}=\overline{\mathcal E_+/\mathcal N_{\rm OS}}\).  The class of
the empty insertion is \(\vac_{\rm OS}\), and finite positive-time insertions
are dense by construction.

Step 2 constructs the positive-energy time evolution.  For \(s\ge0\),
Euclidean time translation by \(s\) maps \(\mathcal E_+\) to itself.  The
semigroup law, OS symmetry, and polynomial matrix-element growth imply
contractivity by Proposition~\ref{prop:os-semigroup-contraction-iteration};
strong continuity is the hypothesis
\eqref{eq:os-time-strong-continuity}.  Hence the descended operators
\(T_E(s)\) form a strongly continuous semigroup of self-adjoint contractions
on \(\Hilb_{\rm OS}\).  The self-adjoint contraction-semigroup theorem gives
a nonnegative self-adjoint generator
\[
  H,\qquad T_E(s)=e^{-sH},\qquad \sigma(H)\subset[0,\infty).
\]
Stone's theorem gives Lorentzian time translations
\[
  U_{\rm time}(t)=\ee^{-\ii tH}.
\]
This is the part of the reconstruction theorem that uses reflection
positivity directly; it does not use the flawed separate-Laplace step of the
first OS paper.

Step 3 constructs the spacetime symmetry representation.  Spatial
translations and rotations preserve the positive-time half-space after a
common positive time shift.  Euclidean covariance and the same null-space
argument used in Step 1 therefore give unitary operators on
\(\Hilb_{\rm OS}\).  Proposition~\ref{prop:os-ii-analytic-route-to-boundary-package}
and Theorem~\ref{thm:corrected-os-ii-analytic-theorem} supply the missing
many-variable analytic continuation: Euclidean rotations mixing time and
space continue to real Lorentzian boosts, and the continued matrix elements
obey the polynomial tube estimates required for tempered boundary values.
The resulting unitary representation \(U\) of the connected Poincare group
has joint translation spectrum in the closed forward light cone.  The forward
spectral support follows from \(H\ge0\), spatial covariance, and the
Lorentzian boost covariance just obtained.

Step 4 constructs the field-valued distributions on the common dense domain.
For a labelled test function \(f\) supported at positive Euclidean time,
define, before taking Lorentzian boundary values,
\[
  \Phi_E(\alpha,f)[F]=[(\alpha,f)\otimes F],
\]
after translating all supports by a common positive Euclidean time if
necessary so that the concatenated configuration is strictly ordered.
Proposition~\ref{prop:os-ordered-field-closability} proves that this
operation descends through the OS null quotient and gives closable operators
on the dense domain \(\mathcal D_{\rm OS}\).  The continuity estimates in the
OS data make all vacuum and finite-domain matrix elements distributions in
the test function.  Their Lorentzian boundary values define the Wightman
distributions and the corresponding operator-valued distributions
\(\widehat\Phi_\alpha\) on \(\mathcal D_{\rm OS}\).

Step 5 verifies the Wightman properties.  Hermiticity of the Schwinger
hierarchy gives the adjoint relation on \(\mathcal D_{\rm OS}\), and
Euclidean covariance gives the Poincare covariance of the reconstructed
fields.  Proposition~\ref{prop:os-boundary-package-consequences} gives
temperedness, the edge-of-the-wedge/Jost-domain proof of local graded
commutativity at spacelike separation, and the reverse continuation back to
the original ordered Euclidean Schwinger functions away from the partial
diagonals.  The restriction to noncoincident Euclidean points is essential:
contact extensions on the diagonals are part of the declared Schwinger
hierarchy and are not recovered from off-diagonal analytic continuation
alone.
\end{proof}

\begin{proposition}[Uniqueness of OS reconstruction]
\label{prop:os-reconstruction-uniqueness}
Two OS reconstructions from the same Schwinger hierarchy are unitarily
equivalent on the reconstructed Wightman field presentation.
\end{proposition}

\begin{proof}
Let
\[
  (\Hilb_1,\vac_1,U_1,\mathcal D_1,\widehat\Phi^{(1)}_\alpha),
  \qquad
  (\Hilb_2,\vac_2,U_2,\mathcal D_2,\widehat\Phi^{(2)}_\alpha)
\]
be two reconstructions from the same hierarchy.  Both start from the same
positive-time vector space \(\mathcal E_+\) and the same OS sesquilinear
form, hence from the same null space \(\mathcal N\).  On the dense quotient
define
\[
  V_0[F]_1=[F]_2 .
\]
This is well defined because \([F]_1=0\) iff
\(\langle F,F\rangle_{\rm OS}=0\) iff \([F]_2=0\).  It is isometric because
both inner products are the same OS form:
\[
  \langle V_0[F]_1,V_0[G]_1\rangle_{\Hilb_2}
  =
  \langle F,G\rangle_{\rm OS}
  =
  \langle [F]_1,[G]_1\rangle_{\Hilb_1}.
\]
Since the quotient is dense in both completions, \(V_0\) extends uniquely to a
unitary \(V:\Hilb_1\to\Hilb_2\).  The empty insertion is mapped to the empty
insertion, so \(V\vac_1=\vac_2\).  Euclidean translations, rotations, and the
continued Poincare transformations are all defined on the dense quotient by
the same operations on test sequences, hence
\[
  VU_1(g)=U_2(g)V
\]
on a dense domain and therefore on the Hilbert spaces.  The same argument
applies to field insertions:
\[
  V\widehat\Phi^{(1)}_\alpha(f)[F]_1
  =
  V[(\alpha,f),F]_1
  =
  [(\alpha,f),F]_2
  =
  \widehat\Phi^{(2)}_\alpha(f)V[F]_1,
\]
with equality on the common invariant domain.  Thus the two reconstructions
are unitarily equivalent as Wightman field presentations.
\end{proof}

\begin{corollary}[path-integral Schwinger limits as Wightman data]
\label{cor:path-integral-schwinger-limit-to-wightman}
Let \(\{\mathcal E_\Lambda,\langle\,\cdot\,\rangle_\Lambda\}\) be a regulated
Euclidean path-integral presentation, where \(\Lambda\) denotes the regulator
and \(\langle\,\cdot\,\rangle_\Lambda\) is the regulated expectation
functional on a declared algebra of Euclidean field insertions.  Assume that
for every \(n\) and every labelled test insertion \(f_n\) in the chosen test
class the regulated moments
\[
  S_{n,\Lambda}(f_n)
  :=
  \langle f_n(\Phi_E,\ldots,\Phi_E)\rangle_\Lambda
\]
converge to a distributional limit \(S_n(f_n)\), and that the limiting
hierarchy \(\{S_n\}_{n\ge0}\) satisfies the OS-admissibility hypotheses of
Definition~\ref{def:os-admissible-euclidean-data}.  Then the continuum path-integral
presentation determines a Wightman field presentation by applying
Theorem~\ref{thm:os-reconstruction} to the limiting Schwinger hierarchy.
If the limiting hierarchy also satisfies the Euclidean cluster condition in
Corollary~\ref{cor:os-reconstruction-vacuum-sector}, the reconstructed
zero-momentum subspace is \(\mathbb C\vac_{\rm OS}\).
\end{corollary}

\begin{proof}
The convergence assumption defines a family
\(\{S_n\}_{n\ge0}\) of continuous linear functionals on the declared
test-function spaces; \(S_0=1\) is the normalization of the regulated
expectation functional after taking the limit.  The OS-admissibility
assumption supplies Euclidean covariance, permutation symmetry, Hermiticity,
reflection positivity, and the OS-II semigroup and linear-growth estimates used in
Theorem~\ref{thm:os-reconstruction}.  Applying that theorem constructs
\[
  (\Hilb_{\rm OS},\vac_{\rm OS},U,\mathcal D_{\rm OS},
  \{\widehat\Phi_\alpha\})
\]
whose Wightman boundary values have the limiting Schwinger hierarchy as their
Euclidean continuation at noncoincident points.  The last sentence is exactly
Corollary~\ref{cor:os-reconstruction-vacuum-sector}.
\end{proof}

\begin{remark}[Constructive status of OS inputs]
\label{rem:os-constructive-status-inputs}
OS reconstruction is a bridge theorem.  It turns a Schwinger hierarchy with
reflection positivity, Euclidean covariance, permutation symmetry,
Hermiticity, clustering when desired, and the OS-II analytic bounds into a
Lorentzian Wightman presentation.  It does not by itself construct the
Schwinger hierarchy.  Completed constructions verifying such inputs are
model-dependent and concentrated in low-dimensional regimes, including
\(P(\phi)_2\), massive \(\phi^4_3\), selected two-dimensional
scalar--fermion models, and two-dimensional Yang--Mills in holonomy or
heat-kernel formulations.  Standard reflection-positive
Ising/\(\lambda\phi^4\) routes in \(D\ge4\) are constrained by triviality
theorems, while four-dimensional Yang--Mills, QCD-like theories, and the
full Standard Model remain open as complete constructive continuum
theorems.  Chapter~\ref{ch:constructive-status-routes} gives the status table
and route-by-route discussion.
\end{remark}

\begin{corollary}[OS reconstruction with and without clustering]
\label{cor:os-reconstruction-vacuum-sector}
Let \(P_0^{\rm OS}\) be the joint spectral projection of the reconstructed
Lorentzian translation representation onto zero momentum.  The reconstruction
theorem by itself produces an invariant cyclic vector \(\vac_{\rm OS}\) and a
positive-energy representation, but it leaves the dimension of
\(P_0^{\rm OS}\Hilb_{\rm OS}\) to the Schwinger hierarchy.  Thus, in the
absence of a cluster condition, the reconstructed representation may contain
several translation-invariant vacuum vectors.

If the Euclidean hierarchy clusters, then for finite polynomial Euclidean
field insertions \(A_E\) and \(B_E\) separated by a large spatial Euclidean
translation,
\[
  \lim_{\lambda\to+\infty}
  \langle A_E\,B_E(\lambda a)\rangle_E
  =
  \langle A_E\rangle_E\,\langle B_E\rangle_E,
  \qquad a\in\mathbb R^{D-1},\ a\ne0,
\]
then its Lorentzian reconstruction satisfies the Wightman weak cluster
property.  Consequently
\[
  P_0^{\rm OS}\Hilb_{\rm OS}=\mathbb C\vac_{\rm OS}
\]
by Theorem~\ref{thm:wightman-cluster-vacuum-uniqueness}.  Conversely, a
one-dimensional reconstructed zero-momentum subspace gives Wightman
clustering, and its restriction back to noncoincident Euclidean points gives
the corresponding Euclidean cluster property.
\end{corollary}

When \(P_0^{\rm OS}\Hilb_{\rm OS}\) has dimension larger than one, the
Schwinger functional is a translation-invariant state that is not extremal. Its
central decomposition is a direct integral, or in finite mixtures a convex
sum, of pure vacuum sectors.  OS positivity and covariance still reconstruct a
valid Wightman presentation, while clustering is the additional condition that
selects a single pure infinite-volume phase.

The theorem has three distinct pieces.  First, reflection positivity constructs
the Hilbert space, vacuum vector, and positive Hamiltonian by the positive-time
quotient.  Second, Euclidean covariance and the OS regularity assumptions
construct the Lorentzian Poincare representation and its spectrum condition.
Third, growth and analyticity assumptions control the distributional boundary
values and produce Wightman functions in the intended test-function class.  In
the tempered case, this target class is \(\mathcal S'(M^n)\).  In larger
analytic frameworks, such as hyperfunction versions of reconstruction, the
target distribution class is correspondingly larger and the growth hypotheses
are changed.

The uniqueness statement is internal to the specified Euclidean hierarchy,
label space, reflection map, and test-function class.  Two reconstructions of
the same OS-admissible data are related by a unitary equivalence preserving the
vacuum, the reconstructed Poincare representation, and the smeared fields.
Equivalence of two different Euclidean presentations, or of the local nets
generated after taking closures and affiliations, is an additional comparison
problem.

\section{Two-Point Spectral Positivity}

For a scalar field, OS positivity gives a direct Euclidean form of spectral
positivity.  Let \(f(\mathbf x)\) be a spatial test function and let
\(\Phi_E(\tau,f)\) denote the reconstructed Euclidean field smeared at
positive time \(\tau\).  The notation \(\Phi_E(0,f)\) below denotes the
boundary value from positive Euclidean time when this domain limit exists.
For \(\tau>0\), time translation in the reconstructed Hilbert space gives
\[
  \langle \vac_{\rm OS},
  \Phi_E(\tau,f)\Phi_E(0,g)\vac_{\rm OS}\rangle_{\rm OS}
  =
  \langle \Phi_E(0,f)^\dagger\vac_{\rm OS},
  \ee^{-\tau H}\Phi_E(0,g)\vac_{\rm OS}\rangle_{\rm OS}.
\]
Applying the spectral theorem to the nonnegative operator \(H\), this is
\[
  \int_{[0,\infty)} \ee^{-\tau E}\,
  \dd\nu_{f,g}(E),
\]
where
\[
  \nu_{f,g}(\Delta)
  =
  \langle \Phi_E(0,f)^\dagger\vac_{\rm OS},
  E_H(\Delta)\Phi_E(0,g)\vac_{\rm OS}\rangle_{\rm OS}
\]
and \(E_H\) is the spectral measure of \(H\).  For \(f=g\), the measure is
positive.  Including spatial translations refines this to the joint spectral
measure of \((H,\mathbf P)\).  Lorentz covariance and the spectrum condition
then repackage the scalar two-point measure by invariant mass, giving the
K\"all\'en--Lehmann representation used in Chapter~\ref{ch:kallen-lehmann}.

This calculation identifies the precise bridge: reflection positivity builds
the Hilbert-space inner product, the Euclidean time semigroup gives the
nonnegative Hamiltonian, and the spectral theorem supplies the positive
energy-momentum measure.

\section{Euclidean Measures and Generating Functionals}

Suppose a Euclidean theory is given by a probability measure \(\mu\) on a
specified measurable space of scalar distributions \(\varphi\), with
generating functional
\[
  Z_E[J]=\int \exp\left(\int_E J(x)\varphi(x)\,\dd^Dx\right)\,\dd\mu(\varphi)
\]
for real test functions \(J\) in a neighborhood of the origin.  The Schwinger
functions are moments:
\[
  S_n(x_1,\ldots,x_n)
  =
  \int \varphi(x_1)\cdots\varphi(x_n)\,\dd\mu(\varphi).
\]
Reflection positivity is then a property of the measure together with the
chosen time reflection.  It is stable under lattice approximations only when
the regulator action, observable insertions, and continuum limit preserve the
positive-time quadratic form.

\begin{remark}
Ordinary positivity of the probability measure \(\mu\) means
\(\int |F(\varphi)|^2\,\dd\mu(\varphi)\ge0\) for measurable functions \(F\).
OS positivity is the reflected statement
\(\int \overline{F(\theta\varphi)}F(\varphi)\,\dd\mu(\varphi)\ge0\) for
positive-time cylinder functions \(F\), with the label adjunction included.
It is this reflected positivity that becomes the Lorentzian Hilbert-space
inner product.
\end{remark}

\section{Reflection Positivity for Generating Functionals}

When the Euclidean moments are encoded by \(Z_E[J]\), reflection positivity
has an equivalent source form.  Let \(J_r\) be finitely many sources supported
in positive Euclidean time and let \(c_r\in\mathbb C\).  Define the reflected
source by
\[
  (\Theta J_r)(x)=\overline{J_r(\theta x)}
\]
with the field adjunction included when the source carries labels.  The
generating functional is reflection positive when
\begin{equation}
  \sum_{r,s}\overline c_r c_s\,
  Z_E[\Theta J_r+J_s]\ge0 .
  \label{eq:os-generating-functional-positivity}
\end{equation}
Expanding \(Z_E\) in powers of the sources recovers
\eqref{eq:os-reflection-positivity} for all Schwinger functions.  Conversely,
if the moment expansion is valid in a neighborhood of the origin, the
Schwinger-function inequalities imply
\eqref{eq:os-generating-functional-positivity}.

For a Gaussian scalar measure with covariance \(C(x-y)\), the criterion
reduces to positivity of the kernel reflected across \(\tau=0\):
\[
  \int_{\tau_x>0}\dd^Dx
  \int_{\tau_y>0}\dd^Dy\,
  \overline{f(x)}\,C(\theta x-y)\,f(y)\ge0
\]
for all positive-time test functions \(f\).  The free massive covariance
\[
  C(x-y)=
  \int\frac{\dd^Dk_E}{(2\pi)^D}
  \frac{\ee^{\ii k_E\cdot(x-y)}}{k_E^2+m^2}
\]
satisfies this condition because its Fourier representation in spatial
momentum is
\[
  C(\tau,\mathbf x)
  =
  \int\frac{\dd^{D-1}\mathbf p}{(2\pi)^{D-1}}
  \frac{\ee^{\ii\mathbf p\cdot\mathbf x}
        \ee^{-\omega_{\mathbf p}|\tau|}}
       {2\omega_{\mathbf p}},
  \qquad
  \omega_{\mathbf p}=\sqrt{\mathbf p^2+m^2},
\]
so the reflected quadratic form is an integral of absolute squares with
positive weight.  This computation is the Euclidean shadow of the positive
one-particle Hilbert space of the free scalar field.

\begin{figure}[H]
  \centering
  \resizebox{0.86\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      box/.style={draw,rounded corners=4pt,line width=0.8pt,fill=orange!7,
        minimum width=3.1cm,minimum height=0.85cm,align=center},
      arrow/.style={-{Latex[length=2mm]},line width=0.85pt}
    }
    \node[box] (measure) at (-4.1,0.9) {Euclidean measure\\\(\mu\)};
    \node[box] (moments) at (0,0.9) {moments\\\(S_n\)};
    \node[box] (os) at (4.1,0.9) {OS-positive\\Schwinger data};
    \node[box,fill=green!7] (wightman) at (0,-1.15)
      {Lorentzian Wightman\\boundary values};
    \draw[arrow] (measure)--(moments);
    \draw[arrow] (moments)--(os)
      node[midway,above=5pt,fill=white,inner sep=1.2pt] {verify axioms};
    \draw[arrow] (os)--(wightman);
    \draw[arrow] (moments)--(wightman)
      node[midway,left] {analytic continuation};
  \end{tikzpicture}
  }
  \caption{A Euclidean measure yields Lorentzian Wightman data through the
  verified OS positivity, covariance, and growth requirements.}
  \label{fig:euclidean-measure-os-data}
\end{figure}

\section{Relation To Wick Rotation}

For time-ordered Green functions in perturbation theory, Wick rotation was
used earlier as a contour and boundary-value operation.  OS reconstruction is
the nonperturbative counterpart at the level of complete Euclidean
correlation data.  It supplies the Hilbert space, Hamiltonian, vacuum, and
Lorentzian Wightman functions from Euclidean data satisfying precise
positivity and growth conditions.

The framework identifies the Euclidean information that reconstructs a
Lorentzian theory: a full compatible hierarchy of Schwinger distributions
together with covariance, reflection positivity, growth, regularity, and
normalization conditions.
