\chapter{Haag--Ruelle Theory and Mathematical Scattering}
\label{ch:volume-iv-haag-ruelle-mathematical-scattering}
\ConventionsLorentzian

The source trace for the massive scattering construction is Haag's asymptotic
condition and Ruelle's proof of scattering states \cite{Haag1958,Ruelle1962}.
The LSZ interface later in the chapter is traced to \cite{LSZ1955}; the text
keeps the Hilbert-space Haag--Ruelle construction logically prior to LSZ
residues.

\section{Vacuum Net and Isolated Mass Shell}

Let \(M=\mathbb M^D\) be Minkowski spacetime, and let
\[
  \mathcal O\longmapsto \mathcal R(\mathcal O)\subset\mathcal B(\Hilb)
\]
be a Poincare-covariant vacuum net of von Neumann algebras.  The standing
data are a Hilbert space \(\Hilb\), a unit vector \(\vac\), and a strongly
continuous unitary representation \(U\) of the proper orthochronous Poincare
group such that
\[
  U(a,\Lambda)\mathcal R(\mathcal O)U(a,\Lambda)^{-1}
  =
  \mathcal R(\Lambda\mathcal O+a),
  \qquad
  U(a,\Lambda)\vac=\vac .
\]
For translations alone we write
\[
  U(a)=U(a,\mathbf 1)=\exp(\ii a_\mu P^\mu),
\]
where \(P^\mu\) are the commuting self-adjoint energy-momentum generators.
The scalar product in Fourier transforms is the same contraction
\(q\cdot x=q_\mu x^\mu\).  This sign convention is part of the standing data:
with the opposite convention for \(U(a)\), every spectral-transfer formula
below would contain \(-\operatorname{supp}\widehat f\) instead of
\(\operatorname{supp}\widehat f\).
The translation generators obey the spectrum condition
\[
  \operatorname{Spec}(P)\subset \overline V_+ .
\]
Let \(E(\Delta)\) denote the joint spectral projection of \(P^\mu\) for a
Borel set \(\Delta\subset\mathbb R^D\).

Fix \(m>0\).  The positive mass shell is
\[
  \Sigma_m^+
  =
  \{p\in\mathbb R^D:p^0=\omega_{\vec p},\ p^2=-m^2\},
  \qquad
  \omega_{\vec p}=\sqrt{\vec p^{\,2}+m^2}.
\]
The one-particle subspace for this stable species is
\[
  \Hilb_1=E(\Sigma_m^+)\Hilb .
\]
This notation carries an existence hypothesis: \(\Hilb_1\ne\{0\}\).  The
projection onto the Borel set \(\Sigma_m^+\) may be zero in a perfectly
consistent Hilbert-space theory, or it may fail to be isolated from continuous
spectral support.  In either case the ordinary massive Haag--Ruelle theorem
does not produce scattering states for that putative species.  The relation
between a one-particle shell and an atom in a K\"all\'en--Lehmann measure was
spelled out in
Remark~\ref{rem:kallen-lehmann-mass-shell-existence-status}.
The Haag--Ruelle construction assumes that the shell is isolated in the
energy-momentum spectrum: there is an open neighborhood \(N\) of
\(\Sigma_m^+\) such that
\[
  \operatorname{Spec}(P)\cap N=\Sigma_m^+ .
\]
This isolation is the spectral form of stability.  The formulas below are
written for a single bosonic species.  If several stable species, spin
multiplets, or fermionic particles are present, \(\Hilb_1\) is replaced by the
appropriate direct sum of isolated one-particle representations and
\(\mathcal F_s(\Hilb_1)\) by the corresponding graded asymptotic Fock space,
with the Bose/Fermi grading supplied in the positive-metric Wightman-field
setting by Theorem~\ref{thm:wightman-spin-statistics}.

For a scalar multiplicity-one species, the one-particle Hilbert space is
realized by the mass-shell spectral representation
\[
  W_1:L^2(\Sigma_m^+,\dd\mu_m)\overset{\sim}{\longrightarrow}\Hilb_1,
  \qquad
  \dd\mu_m(p)=\frac{\dd^{D-1}\vec p}{(2\pi)^{D-1}2\omega_{\vec p}},
\]
with
\[
  W_1 f=\int_{\Sigma_m^+}\dd\mu_m(p)\,f(p)|p\rangle,
  \qquad
  U(a,\mathbf 1)W_1f
  =
  W_1\bigl(p\mapsto \ee^{\ii a_\mu p^\mu}f(p)\bigr).
\]
The bosonic asymptotic Fock space is then the Hilbert direct sum
\[
  \mathcal F_s(\Hilb_1)
  \simeq
  \mathbb C\Omega_F
  \oplus
  \bigoplus_{n\ge1}
  L^2_{\mathrm{sym}}\bigl((\Sigma_m^+)^n,\dd\mu_m^{\otimes n}\bigr).
\]
Thus the wave operators below start from a constructed free many-particle
Hilbert space whose momentum labels are coordinates in the spectral model.

\section{Almost-Local Operators and Spectral Transfer}

Almost-locality is the norm-approximation condition of
Definition~\ref{def:almost-local-bounded-operator}: an operator
\(B\in\mathcal B(\Hilb)\) is approximated by bounded local operators \(B_R\)
localized in double cones of radius \(R\) such that, for every \(N\ge1\), some
\(C_N<\infty\) obeys
\[
  \|B-B_R\|\le C_N R^{-N}.
\]
Almost locality is the operator-algebraic substitute for using smeared local
fields whose support is not strictly compact after energy-momentum
filtering.

For a bounded local operator \(A\), set
\[
  A(x)=U(x)AU(x)^{-1},\qquad x\in\mathbb R^D .
\]
For \(f\in\mathcal S(\mathbb R^D)\), define
\[
  A(f)=\int\dd^D x\,f(x)A(x),
  \qquad
  \widehat f(q)=\int\dd^D x\,f(x)\ee^{\ii q\cdot x}.
\]
With this convention, the energy-momentum transfer of \(A(f)\) is contained
in \(\operatorname{supp}\widehat f\):
\begin{equation}
  E(\Delta_1)A(f)E(\Delta_2)=0
  \quad\text{if}\quad
  \Delta_1\cap
  \bigl(\Delta_2+\operatorname{supp}\widehat f\bigr)=\emptyset .
  \label{eq:haag-ruelle-spectral-transfer}
\end{equation}
Indeed, by the joint spectral theorem,
\[
  E(\dd p)\,A(f)\,E(\dd r)
  =
  \left(\int \dd^D x\,f(x)\ee^{\ii x_\mu(p^\mu-r^\mu)}\right)
  E(\dd p)\,A\,E(\dd r)
  =
  \widehat f(p-r)\,E(\dd p)\,A\,E(\dd r).
\]
After integrating \(p\in\Delta_1\) and \(r\in\Delta_2\), the operator can be
nonzero only if \(p-r\in\operatorname{supp}\widehat f\), equivalently
\(\Delta_1\cap(\Delta_2+\operatorname{supp}\widehat f)\ne\emptyset\).
Equation \eqref{eq:haag-ruelle-spectral-transfer} is the precise meaning of
selecting a particle shell by smearing in spacetime.

Choose a compact set \(K\subset\Sigma_m^+\).  Let \(\chi\in\mathcal S(\mathbb R^D)\)
have \(\operatorname{supp}\widehat\chi\subset N\), with
\(\widehat\chi=1\) on a smaller neighborhood of \(K\).  For a local \(A\),
set
\[
  B=A(\chi).
\]
Then \(B\) is almost local.  Since the vacuum has momentum \(0\),
\[
  B\vac=E(\operatorname{supp}\widehat\chi)B\vac=P_1B\vac,
  \qquad P_1=E(\Sigma_m^+),
\]
provided the support of \(\widehat\chi\) meets the spectrum only on the
isolated shell.  If \(-\operatorname{supp}\widehat\chi\) is disjoint from the
forward spectrum and \(\operatorname{supp}\widehat\chi\) avoids \(0\), then
\[
  B^*\vac=0 .
\]
An almost-local operator satisfying these spectral support properties and
\(B\vac\ne0\) is a regular Haag--Ruelle creator for the shell.

Figure~\ref{fig:haag-ruelle-spectral-transfer} displays the spectral-transfer
condition for such a creator: the Fourier support meets the spectrum only on
the isolated positive mass shell and avoids the vacuum and negative-energy
regions.

\begin{figure}[H]
  \centering
  \resizebox{0.78\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \draw[-{Latex[length=2mm]},line width=0.75pt] (-3.2,0)--(3.45,0)
      node[right] {\(\vec p\)};
    \draw[-{Latex[length=2mm]},line width=0.75pt] (0,-2.35)--(0,3.2)
      node[above] {\(p^0\)};
    \draw[gray!50,dashed,line width=0.8pt] (-2.9,2.9)--(0,0)--(2.9,2.9);
    \node[gray!60!black] at (2.65,2.45) {\(\overline V_+\)};
    \draw[blue!70!black,line width=1.1pt]
      plot[smooth,domain=-2.25:2.25,samples=90]
      (\x,{sqrt(\x*\x+0.95)});
    \node[blue!70!black] at (-2.25,2.65) {\(\Sigma_m^+\)};
    \fill[green!14,draw=green!45!black,line width=0.9pt]
      (1.12,1.55) ellipse (0.72 and 0.43);
    \node[green!45!black,align=center] at (1.12,0.72)
      {\(\operatorname{supp}\widehat\chi\subset N\)};
    \fill[orange!15,draw=orange!85!black,line width=0.9pt]
      (-1.12,-1.55) ellipse (0.72 and 0.43);
    \node[orange!85!black,align=center] at (-1.42,-0.72)
      {transfer of \(B^*\)};
    \fill[black] (0,0) circle (1.7pt);
    \node[below right] at (0,0) {\(\vac\)};
  \end{tikzpicture}
  }
  \caption{A spacetime smearing with \(\chi\) restricts energy-momentum
  transfer to \(\operatorname{supp}\widehat\chi\).  The support is chosen
  inside the isolating neighborhood of the positive mass shell, while the
  reflected support controls the adjoint creator.}
  \label{fig:haag-ruelle-spectral-transfer}
\end{figure}

\section{Wave Packets and Velocity Localization}

Let \(B\) be a regular creator.  For
\[
  h\in C_c^\infty(\mathbb R^{D-1}),
\]
define the positive-energy Klein--Gordon packet
\[
  h_t(\vec x)
  =
  \int\frac{\dd^{D-1}\vec p}{(2\pi)^{D-1}}\,
  h(\vec p)\,
  \ee^{\ii\omega_{\vec p}t-\ii\vec p\cdot\vec x}
\]
and the time-dependent operator
\[
  B_t(h)=\int\dd^{D-1}\vec x\,h_t(\vec x)B(t,\vec x).
\]
The one-particle vector
\[
  \psi_h=P_1B_t(h)\vac
\]
is independent of \(t\); the phase in \(h_t\) cancels the free time evolution
on the mass shell.

The velocity support of \(h\) is
\[
  V(h)=
  \left\{
    \frac{\vec p}{\omega_{\vec p}}:
    \vec p\in\operatorname{supp}h
  \right\}.
\]
Stationary phase says that \(h_t(\vec x)\) is rapidly decreasing away from
the equal-time tube
\[
  \vec x\in t\,V(h)+O(|t|^0).
\]
Thus packets with disjoint velocity supports become spacelike separated at
large \(|t|\).  Combining this geometric separation with almost locality
gives, for every \(N\ge1\),
\begin{equation}
  \|[B_{i,t}(h_i),B_{j,t}(h_j)]\|
  \le C_N(1+|t|)^{-N}
  \label{eq:algebraic-haag-ruelle-commutator}
\end{equation}
whenever \(V(h_i)\cap V(h_j)=\emptyset\).

\section{Haag--Ruelle Limits}

Let \(B_1,\ldots,B_n\) be regular creators and let
\[
  h_1,\ldots,h_n\in C_c^\infty(\mathbb R^{D-1})
\]
have pairwise disjoint velocity supports.  Define
\[
  \Psi_t=
  B_{1,t}(h_1)\cdots B_{n,t}(h_n)\vac .
\]

\begin{theorem}[Haag--Ruelle scattering for a vacuum net]
\label{thm:algebraic-haag-ruelle-scattering}
Assume locality, Poincare covariance, the spectrum condition, a unique
vacuum, an isolated massive one-particle shell, and enough regular creators
to generate a dense subspace of \(\Hilb_1\).  Then the limits
\[
  \Psi_{\mathrm{out}}=\lim_{t\to+\infty}\Psi_t,
  \qquad
  \Psi_{\mathrm{in}}=\lim_{t\to-\infty}\Psi_t
\]
exist in Hilbert norm.  They depend only on the one-particle vectors
\(\psi_{h_j}=P_1B_{j,t}(h_j)\vac\), and they define isometric wave
operators from the symmetric Fock space \(\mathcal F_s(\Hilb_1)\) into
\(\Hilb\).
\end{theorem}

\begin{remark}[Physical scope of the isolated-shell hypothesis]
\label{rem:haag-ruelle-isolated-shell-scope}
The nonzero isolated-shell assumption is a spectral theorem about the
particular nonperturbative model, not a generic consequence of writing a formal
Lagrangian.  It is satisfied by free massive fields.  It is also available in
the massive stable sectors of constructive low-dimensional scalar models, such
as \(P(\phi)_2\) and massive \(\lambda\phi^4_3\), in regimes where the
Wightman theory, mass gap, and one-particle spectral gap have been established.
For a four-dimensional \(\lambda\phi^4\) Lagrangian treated only
perturbatively, an isolated real pole in the formal two-point function is
evidence for the intended particle interpretation, but it is not a
nonperturbative theorem constructing an isolated shell in a Hilbert-space
spectrum.

A K\"all\'en--Lehmann measure with no atom at \(m^2\), even if its continuous
support starts at \(m^2\), supplies no normalizable one-particle vector of mass
\(m\) for the corresponding local interpolating field.  Haag--Ruelle then has
no external particle of that species to scatter.  The theorem exactly
distinguishes a sharp stable particle from threshold spectral weight.

The hypothesis fails, or is the wrong hypothesis, in several central physical
settings.  In QED with long-range electromagnetic fields, charged sectors are
expected to be infraparticle sectors: the energy-momentum spectrum has
continuous soft-photon dressing attached to the charged state rather than a
sharp Wigner mass hyperboloid.  The photon is massless and belongs to a theory
without the massive spectral gap used by the theorem.  In QCD, colored quark
and gluon shells are not expected to occur in the gauge-invariant observable
Hilbert space; stable hadrons, when present below strong-decay thresholds, are
the candidate isolated shells.  Unstable particles and resonances do not give
external Haag--Ruelle particles, because their spectral singularities are
resonance poles on analytically continued sheets rather than point spectrum on
the physical mass shell.  Each of these cases requires a different asymptotic
framework or a separate spectral theorem.  The corresponding
asymptotic-completeness problems for charged QED and physical QCD are stated
as Open Problems~\ref{op:qed-charged-asymptotic-completeness}
and~\ref{op:qcd-asymptotic-completeness}.
\end{remark}

\subsection{The Cook Estimate and the Fock Inner Product}

We now spell out the proof in a form that separates the analytic estimates
from the algebraic conclusion.  The estimates below are precisely the
Haag--Ruelle estimates supplied by almost locality, spectral isolation, and
stationary phase.  Once they are known, the construction of scattering states
and the Fock inner product is an elementary Hilbert-space argument.

\begin{definition}[Haag--Ruelle estimate package]
\label{def:haag-ruelle-estimate-package}
For a finite family of regular creators \(B_j\) and smooth compactly
supported packets \(h_j\), write
\[
  B_{j,t}=B_{j,t}(h_j),
  \qquad
  \psi_j=B_{j,t}\vac\in\Hilb_1 .
\]
The family is said to satisfy the Haag--Ruelle estimate package on
\(|t|\ge T_0\) if the following hold.
\begin{enumerate}
  \item The vectors \(B_{j,t}\vac=\psi_j\) are independent of \(t\), and
  \(B_{j,t}^*\vac=0\).
  \item For every \(N\ge1\) and every pair with disjoint velocity supports,
  \begin{equation}
    \|[B_{i,t}^{\#},B_{j,t}^{\#}]\|
    \le C_N(1+|t|)^{-N},
    \qquad B^\#\in\{B,B^*\}.
    \label{eq:hr-estimate-package-commutator}
  \end{equation}
  The same estimate holds on the finite-energy vectors obtained from the
  family when one entry is replaced by its strong time derivative.  This
  derivative is the derivative of the smooth translation-smearing parameter;
  no norm differentiability of an unsmeared local operator is assumed.
  \item For every pair of regular creators \(B,C\) with packets \(h,g\),
  \begin{equation}
    B_t(h)^*C_t(g)\vac
    =
    \langle B_t(h)\vac,C_t(g)\vac\rangle\,\vac
    +
    r_t(B,h;C,g),
    \label{eq:hr-estimate-package-contraction}
  \end{equation}
  where for every \(N\ge1\),
  \(\|r_t(B,h;C,g)\|\le C_N(1+|t|)^{-N}\).
  \item For two velocity-separated lists
  \(B_1,\ldots,B_n\) and \(C_1,\ldots,C_m\), the scalar products obey the
  recursive contraction estimate
  \begin{equation}
  \begin{aligned}
  &
  \left\langle
    B_{1,t}\cdots B_{n,t}\vac,
    C_{1,t}\cdots C_{m,t}\vac
  \right\rangle
  \\
  &\quad =
  \sum_{\ell=1}^{m}
  \langle\psi_1,\varphi_\ell\rangle
  \left\langle
    B_{2,t}\cdots B_{n,t}\vac,
    C_{1,t}\cdots \widehat{C_{\ell,t}}\cdots C_{m,t}\vac
  \right\rangle
  +O_N((1+|t|)^{-N})
  \end{aligned}
  \label{eq:hr-estimate-package-recursive-contraction}
  \end{equation}
  for every \(N\ge1\), with the convention that the sum is empty when
  \(m=0\).
  \item For each velocity-separated product
  \(\Psi_t=B_{1,t}\cdots B_{n,t}\vac\), there is an integrable nonnegative
  function \(R(t)\) on \([T_0,\infty)\) and on
  \((-\infty,-T_0]\) such that
  \begin{equation}
    \left\|\frac{\dd}{\dd t}\Psi_t\right\|\le R(t)
    \label{eq:hr-estimate-package-cook}
  \end{equation}
  on the finite vector domain under discussion.
\end{enumerate}
\end{definition}

\begin{proposition}[Estimate package implies Haag--Ruelle limits]
\label{prop:estimate-package-implies-haag-ruelle}
Assume the estimate package
Definition~\ref{def:haag-ruelle-estimate-package} for every finite
velocity-separated family of regular creators.  Then the limits in
Theorem~\ref{thm:algebraic-haag-ruelle-scattering} exist, depend only on the
one-particle vectors, and satisfy the bosonic Fock inner product formula
\begin{equation}
\begin{aligned}
&
\left\langle
  \Psi^{\rm out}(\psi_1,\ldots,\psi_n),
  \Phi^{\rm out}(\varphi_1,\ldots,\varphi_m)
\right\rangle
\\
&\qquad
=
\delta_{nm}
\sum_{\sigma\in S_n}
\prod_{j=1}^{n}
\langle\psi_j,\varphi_{\sigma(j)}\rangle_{\Hilb_1}.
\end{aligned}
\label{eq:hr-fock-inner-product-from-estimates}
\end{equation}
The same statement holds for incoming limits.  Hence the wave operators are
isometries from \(\mathcal F_s(\Hilb_1)\) into \(\Hilb\).
\end{proposition}

\begin{proof}
We give the outgoing proof.  The incoming proof is the same with
\(t\to-\infty\).

First, Cook's criterion gives existence.  For \(t'>t\ge T_0\),
\[
  \|\Psi_{t'}-\Psi_t\|
  \le
  \int_t^{t'}\left\|\frac{\dd}{\dd s}\Psi_s\right\|\,\dd s
  \le
  \int_t^{t'}R(s)\,\dd s .
\]
Since \(R\in L^1([T_0,\infty))\), the right side tends to zero as
\(t,t'\to+\infty\).  Thus \(\Psi_t\) is Cauchy in the Hilbert norm and has an
outgoing limit.

Second, compute the limiting scalar products.  Let
\[
  \Psi_t=B_{1,t}\cdots B_{n,t}\vac,
  \qquad
  \Phi_t=C_{1,t}\cdots C_{m,t}\vac ,
\]
with both lists velocity-separated.  Then
\[
  \langle\Psi_t,\Phi_t\rangle
  =
  \langle\vac,
  B_{n,t}^*\cdots B_{1,t}^*
  C_{1,t}\cdots C_{m,t}\vac\rangle .
\]
The recursive contraction estimate
\eqref{eq:hr-estimate-package-recursive-contraction} gives
\[
  \langle\Psi_t,\Phi_t\rangle
  =
  \sum_{\ell=1}^{m}
  \langle\psi_1,\varphi_\ell\rangle\,
  \langle
    B_{2,t}\cdots B_{n,t}\vac,
    C_{1,t}\cdots \widehat{C_{\ell,t}}\cdots C_{m,t}\vac
  \rangle
  +O_N(t^{-N}).
\]
Induction on \(n+m\), with the base cases following from
\(\langle\vac,\vac\rangle=1\) and \(B_t^*\vac=0\), gives
\eqref{eq:hr-fock-inner-product-from-estimates}.  If \(n\ne m\), the
induction eventually leaves an annihilator acting on the vacuum, and the
limit is zero.

The scalar product formula shows at once that changing an interpolating
creator while keeping the one-particle vector fixed changes the limiting
state by a vector of zero norm.  It also identifies the norm of every finite
linear combination of outgoing limits with the norm of the corresponding
finite symmetric tensor in \(\mathcal F_s(\Hilb_1)\).  Therefore the map from
the algebraic velocity-separated Fock domain to \(\Hilb\) is isometric and
extends uniquely to the Fock completion.  Density of regular creators in
\(\Hilb_1\) gives the stated domain of the wave operator.
\end{proof}

\begin{lemma}[Stationary-phase velocity localization]
\label{lem:hr-stationary-phase-velocity-localization}
Let \(h\in C_c^\infty(\mathbb R^{D-1})\) and let
\[
  h_t(\vec x)
  =
  \int\frac{\dd^{D-1}\vec p}{(2\pi)^{D-1}}\,
  h(\vec p)\,
  \ee^{\ii\omega_{\vec p}t-\ii\vec p\cdot\vec x}.
\]
Let \(V(h)=\{\vec p/\omega_{\vec p}:\vec p\in\operatorname{supp}h\}\).  If
\(U\subset\mathbb R^{D-1}\) is an open neighborhood of \(V(h)\), then for
every \(N\ge1\) there is \(C_N\) such that
\[
  |h_t(\vec x)|\le C_N(1+|t|+|\vec x|)^{-N}
  \qquad
  \text{whenever }\vec x/t\notin U,\quad |t|\ge1 .
\]
\end{lemma}

\begin{proof}
On the compact support of \(h\), the phase is
\[
  \Phi_t(\vec p;\vec x)
  =
  \omega_{\vec p}t-\vec p\cdot\vec x .
\]
Its gradient is
\[
  \nabla_{\vec p}\Phi_t
  =
  t\,\frac{\vec p}{\omega_{\vec p}}-\vec x .
\]
If \(\vec x/t\notin U\), compactness gives
\(|\nabla_{\vec p}\Phi_t|\ge c|t|\) on \(\operatorname{supp}h\) for some
\(c>0\).  Apply repeatedly the integration-by-parts operator
\[
  L
  =
  \frac{1}{\ii|\nabla_{\vec p}\Phi_t|^2}
  \nabla_{\vec p}\Phi_t\cdot\nabla_{\vec p},
  \qquad
  L\ee^{\ii\Phi_t}=\ee^{\ii\Phi_t}.
\]
Each integration by parts gains a factor \(O(|t|^{-1})\), while derivatives
of \(h\) and of the coefficients of \(L\) are bounded on the compact support.
This gives \(O(|t|^{-N})\).  In the region \(|\vec x|\gg |t|\), the same
lower bound is \(c(|t|+|\vec x|)\), giving the displayed estimate.
\end{proof}

\begin{proposition}[Almost locality gives the commutator estimate]
\label{prop:almost-locality-gives-hr-commutator}
Let \(B_i\) and \(B_j\) be almost-local regular creators and let
\(h_i,h_j\) have disjoint velocity supports.  Then
\[
  \|[B_{i,t}(h_i),B_{j,t}(h_j)]\|
  \le C_N(1+|t|)^{-N}
  \qquad (N\ge1),
\]
and the same estimate holds with either operator replaced by its adjoint.
\end{proposition}

\begin{proof}
Choose nested open neighborhoods
\[
  V(h_\ell)\subset U_\ell^0\Subset U_\ell,\qquad \ell=i,j,
\]
so that \(U_i\) and \(U_j\) have positive Euclidean distance
\[
  \delta=\operatorname{dist}(U_i,U_j)>0 .
\]
Let \(\chi_i^t,\chi_j^t\) be smooth cutoffs equal to one on the smaller
equal-time tubes \(\vec x/t\in U_i^0\), \(\vec y/t\in U_j^0\), and supported
in the larger disjoint tubes \(\vec x/t\in U_i\), \(\vec y/t\in U_j\).
Write
\[
  B_{i,t}=B_{i,t}^{\rm near}+B_{i,t}^{\rm tail},
  \qquad
  B_{i,t}^{\rm near}
  =
  \int \dd^{D-1}\vec x\,\chi_i^t(\vec x)h_{i,t}(\vec x)B_i(t,\vec x),
\]
and similarly for \(j\).  The stationary-phase lemma, used with an exponent
larger than \(N+D\), gives an \(L^1\)-tail estimate
\[
  \int_{\vec x/t\notin U_i^0}|h_{i,t}(\vec x)|\,\dd^{D-1}\vec x
  \le C_N(1+|t|)^{-N},
\]
and the same for \(h_j\).  Since translations preserve operator norms,
\[
  \|B_{i,t}^{\rm tail}\|
  \le \|B_i\|
  \int_{\vec x/t\notin U_i}|h_{i,t}(\vec x)|\,\dd^{D-1}\vec x
  =
  O_N(|t|^{-N}) .
\]
Every commutator containing at least one tail factor is therefore
\(O_N(|t|^{-N})\).

It remains to estimate the commutator of the near parts.  For
\(\vec x/t\in U_i\) and \(\vec y/t\in U_j\) the equal-time centers satisfy
\[
  |\vec x-\vec y|\ge \delta |t| .
\]
Choose almost-local approximants \(B_{\ell,R}\) localized in the double cone
of radius \(R\) about the origin and obeying, for arbitrary \(K\),
\[
  \|B_\ell-B_{\ell,R}\|\le C_K R^{-K},
  \qquad \ell=i,j .
\]
Set \(R=\delta |t|/8\).  A point in the translated localization region of
\((B_{i,R})(t,\vec x)\) has time coordinate within \(R\) of \(t\) and spatial
coordinate within \(R\) of \(\vec x\); likewise for \(j\).  Hence the
absolute time separation of two points in the two translated double cones is
at most \(2R\), while their spatial separation is at least
\(\delta |t|-2R\).  With the chosen \(R\),
\[
  \delta |t|-2R=\frac34\delta |t|>2R,
\]
so the two double cones are spacelike separated.  Locality gives
\[
  [(B_{i,R})(t,\vec x),(B_{j,R})(t,\vec y)]=0
\]
for all \(\vec x,\vec y\) in the near supports, and hence the integrated
commutator of the two approximated near parts vanishes.

The error made by replacing \(B_\ell\) by \(B_{\ell,R}\) is controlled in
operator norm.  The near-packet \(L^1\)-norm grows at most polynomially in
\(|t|\): inside a ball of radius \(O(|t|)\) this is the trivial volume bound,
while outside such a ball the same nonstationary-phase integration by parts
gives rapid decay.  Thus, for some fixed power \(M\),
\[
  \|B_{\ell,t}^{\rm near}-(B_{\ell,R})_{t}^{\rm near}\|
  \le C(1+|t|)^M R^{-K}.
\]
Since \(K\) is arbitrary in the almost-local approximation, choosing
\(K\) larger than \(M+N\) makes the replacement error \(O_N(|t|^{-N})\).
Combining the near-approximant commutator, the replacement errors, and the
tail estimates proves the displayed commutator bound.  The adjoint case is
identical: \(B_R^*\) is localized in the same double cone as \(B_R\), and
\(\|B^*-B_R^*\|=\|B-B_R\|\), so almost locality is preserved by taking
adjoints.
\end{proof}

\begin{proposition}[Origin of the remaining Haag--Ruelle estimates]
\label{prop:origin-of-remaining-haag-ruelle-estimates}
For regular creators obtained by smooth spectral transfer inside the
isolating neighborhood \(N\), the one-particle contraction estimate
\eqref{eq:hr-estimate-package-contraction}, the recursive contraction
estimate \eqref{eq:hr-estimate-package-recursive-contraction}, and the Cook
derivative estimate \eqref{eq:hr-estimate-package-cook} follow from spectral
isolation of \(\Sigma_m^+\), the transfer relation
\eqref{eq:haag-ruelle-spectral-transfer}, and the commutator estimate of
Proposition~\ref{prop:almost-locality-gives-hr-commutator}.
\end{proposition}

\begin{proof}
The support of the spectral transfer of a regular creator meets the spectrum,
when acting on the vacuum, only on \(\Sigma_m^+\).  Therefore
\(B_t(h)\vac\) is exactly the one-particle vector determined by the mass-shell
packet and is independent of \(t\), while \(B_t(h)^*\vac=0\) because the
reflected transfer support is disjoint from \(\overline V_+\).  Acting with
\(B_t(h)^*C_t(g)\) on the vacuum, spectral transfer permits a vacuum component
and separates all nonvacuum components from the isolated shell by a positive
spectral distance fixed by the supports.  Choose a smooth cutoff \(\rho(P)\)
that is \(0\) on a neighbourhood of the vacuum point and \(1\) on the
separated nonvacuum spectral support.  The oscillatory factors left after
subtracting the vacuum component then have no stationary point on the relevant
compact spectral support.  Repeated integration by parts in the spectral
variables, written equivalently as commuting powers of smooth functions of
\(P\) through the translated smearing, gives the rapidly decreasing remainder
in \eqref{eq:hr-estimate-package-contraction}.  The estimate is a strong
finite-energy estimate; it is not a statement about an operator norm
derivative of a pointlike field.
Iterating this contraction after decomposing packet supports into finitely
many small velocity cells gives
\eqref{eq:hr-estimate-package-recursive-contraction}; terms in which two
uncontracted cells have separated velocity supports are estimated by
Proposition~\ref{prop:almost-locality-gives-hr-commutator}, and the finite
partition is removed by summing the finitely many cells.

For the Cook estimate, differentiate the product on the finite vector domain
spanned by the vectors obtained by applying finitely many translated,
spectrally smeared creators to the vacuum.
The derivative of the \(k\)-th factor, after commuting it to the far right,
annihilates the vacuum because the one-particle packet has been chosen to
solve the free mass-shell evolution.  Every commutation needed to move this
derivative through later factors is controlled by the derivative version of
Proposition~\ref{prop:almost-locality-gives-hr-commutator}; the derivative
version is obtained by applying
Lemma~\ref{lem:hr-stationary-phase-velocity-localization} to the differentiated
packet.  The differentiated packet only multiplies the compact momentum
amplitude by a smooth polynomial expression in \(p\), so it has the same
velocity localization and rapid tube-tail estimates.
The resulting bound is \(C_N(1+|t|)^{-N}\) for arbitrary \(N\), hence
integrable.  This is \eqref{eq:hr-estimate-package-cook}.
\end{proof}

Together,
Propositions~\ref{prop:estimate-package-implies-haag-ruelle},
\ref{prop:almost-locality-gives-hr-commutator}, and
\ref{prop:origin-of-remaining-haag-ruelle-estimates} give the ordinary
massive Haag--Ruelle theorem under the displayed vacuum-net, isolated-shell,
and regular-creator hypotheses.  No perturbative pole expansion or
Lagrangian formalism enters this construction.

\begin{proof}[Proof of Theorem~\ref{thm:algebraic-haag-ruelle-scattering}]
Choose regular creators \(B_j\) whose one-particle projections are the desired
wave packets and whose velocity supports are pairwise disjoint after the usual
finite partition of momentum space.  Spectral isolation of the mass shell gives
\(B_{j,t}\vac=P_1B_{j,t}\vac\) independent of \(t\), and excludes
\(B_{j,t}^*\vac\) by the support condition against the forward spectrum.
Lemma~\ref{lem:hr-stationary-phase-velocity-localization} and almost locality
give the rapid commutator bound through
Proposition~\ref{prop:almost-locality-gives-hr-commutator}.  The remaining
contraction and Cook bounds are supplied by
Proposition~\ref{prop:origin-of-remaining-haag-ruelle-estimates}; hence the
family satisfies Definition~\ref{def:haag-ruelle-estimate-package}.  Applying
Proposition~\ref{prop:estimate-package-implies-haag-ruelle} gives the outgoing
and incoming Hilbert-norm limits and the Fock scalar product.  The scalar
product identifies vectors obtained from different interpolating creators with
the same one-particle projection and extends the wave maps by completion from
the algebraic symmetric Fock domain.  The assumed density of regular creators
in \(\Hilb_1\) gives the stated wave operators on
\(\mathcal F_s(\Hilb_1)\).
\end{proof}

\section{Gauge Charges and Nonlocal Charged Creators}

The ordinary theorem uses local creators \(B\in\Obs(\mathcal O)\) whose
vacuum vector \(B\vac\) has nonzero projection to an isolated one-particle
shell.  In a gauge theory with an unscreened Gauss-law charge, the existence
of such a creator is a further dynamical and localization hypothesis.  It is
false for charged particles if the available local algebra is the
gauge-invariant observable algebra.

\begin{theorem}[Gauss-law obstruction to local charged creators]
\label{thm:gauss-law-obstruction-local-charged-creators}
Let \(\Obs\) be a gauge-invariant local observable net in a vacuum
representation.  Assume that an abelian charge \(Q\) is represented on a
common invariant domain by the large-sphere Gauss-law limit
\[
  Q=\lim_{R\to\infty}Q_R,\qquad
  Q_R=\int_{S_R} n_i E^i\,\dd S ,
\]
and that locality holds between \(\Obs(\mathcal O)\) and electric-field
operators localized spacelike to \(\mathcal O\).  If the vacuum is neutral,
\(Q\vac=0\), then every local gauge-invariant observable \(A\in\Obs(\mathcal O)\)
creates a neutral vector:
\[
  Q\,A\vac=0 .
\]
Consequently a charged one-particle vector cannot be obtained by applying a
bounded-region gauge-invariant observable to the vacuum.
\end{theorem}

\begin{proof}
The surface charge must be read as a limit of smeared operators.  Choose a
smooth function \(\rho_R(r)\) supported in a shell \(R<r<2R\), normalized so
that it represents the outward flux through a large sphere, and define the
smeared Gauss operator \(Q_R(\rho)\) on the common invariant domain.  For
\(R\) larger than the spatial diameter of \(\mathcal O\), the support of the
electric-field smearing is spacelike to \(\mathcal O\) at the chosen time
slice.  Locality of the observable net with the electric-field algebra gives
\[
  [Q_R(\rho),A]=0
\]
on that domain.  The statement is insensitive to the particular smooth
representative of the large sphere: changing \(\rho_R\) inside the exterior
region changes \(Q_R(\rho)\) by an operator localized spacelike to
\(\mathcal O\), plus terms that vanish in the assumed Gauss-law limit.

By hypothesis the strong graph limit of \(Q_R(\rho)\) on vectors generated
from the vacuum by local observables exists and is the charge \(Q\).  Hence
for every vector \(\Phi\) in the common domain,
\[
  \langle \Phi,(QA-AQ)\vac\rangle
  =
  \lim_{R\to\infty}\langle \Phi,[Q_R(\rho),A]\vac\rangle
  =
  0 .
\]
Thus \(QA\vac=AQ\vac\) as a vector identity on the charge domain.  Since the
vacuum is neutral,
\[
  Q A\vac=AQ\vac=0 .
\]
If \(\psi\) has charge \(q\ne0\), so that \(Q\psi=q\psi\), it cannot equal
\(A\vac\) for such an \(A\).
\end{proof}

This obstruction concerns localization and Gauss-law charge.  It is distinct
from the infraparticle obstruction in massless QED.  The latter states that a
charged sector with long-range electromagnetic flux does not contain an
ordinary sharp Wigner mass-shell vector.  The present theorem applies before
asking whether the charged spectral distribution is an isolated shell or a
soft-photon continuum: a bounded gauge-invariant observable cannot change the
surface charge at infinity.

\paragraph{Almost-local closure.}
Under the hypotheses of
Theorem~\ref{thm:gauss-law-obstruction-local-charged-creators}, assume in
addition that the charge operator \(Q\) is closed on a domain containing the
vectors generated from \(\vac\) by local gauge-invariant observables.  Let
\(B\in\mathcal B(\Hilb)\) be almost local with respect to the observable net:
there are \(B_R\in\Obs(\mathcal O_R)\), with
\(\mathcal O_R\) a cofinal family of bounded regions, such that
\[
  \|B-B_R\|\longrightarrow0
  \qquad (R\to\infty).
\]
Then \(B\vac\) is neutral:
\[
  B\vac\in\operatorname{Dom}(Q),
  \qquad QB\vac=0 .
\]
Consequently a vector with nonzero Gauss charge cannot be produced from the
vacuum by an almost-local operator belonging to the gauge-invariant
observable net.

For each \(R\), Theorem~\ref{thm:gauss-law-obstruction-local-charged-creators}
applied to the local observable \(B_R\) gives
\[
  B_R\vac\in\operatorname{Dom}(Q),
  \qquad Q B_R\vac=0 .
\]
The norm convergence \(B_R\to B\) implies \(B_R\vac\to B\vac\), while
\(QB_R\vac\to0\).  Closedness of \(Q\) therefore gives
\[
  B\vac\in\operatorname{Dom}(Q),
  \qquad QB\vac=0 .
\]
If \(B\vac\) had a nonzero component in a charge-\(q\) spectral subspace,
\(q\ne0\), the self-adjoint spectral calculus for \(Q\) would give
\(QB\vac\ne0\) on that component.  Hence no such component is present.

This almost-local extension is the exact point at which the ordinary
Haag--Ruelle hypothesis fails.  The obstruction is structural: the noncompact
tail is forced by the charge superselection law.  If an operator changes the
Gauss-law flux at infinity, it cannot be approximated in norm, even rapidly,
by bounded
gauge-invariant observables localized in bounded regions.

Charged fields can be introduced in a larger field algebra, but they are not
gauge-invariant local observables.  A gauge-invariant charged insertion must
carry flux to infinity.  In an abelian convention
\[
  A\longmapsto A+\dd\alpha,\qquad
  \psi_q(x)\longmapsto \ee^{\ii q\alpha(x)}\psi_q(x),
\]
a path \(\gamma\) from \(x\) to spatial infinity gives the formal dressing
\begin{equation}
  \Psi_{q,\gamma}(x)
  =
  \psi_q(x)\,
  \exp\!\left\{\ii q\int_\gamma A\right\}.
  \label{eq:charged-wilson-line-dressed-field}
\end{equation}
For gauge functions with \(\alpha(\infty)=0\), the exponential transforms by
\(\ee^{-\ii q\alpha(x)}\), so \(\Psi_{q,\gamma}(x)\) is gauge invariant.  For
a gauge function with nonzero boundary value, it transforms by the boundary
phase and therefore carries the global charge.  The price is noncompact
localization: the localization region includes the curve \(\gamma\) reaching
infinity, or a Coulombic dressing with equivalent asymptotic flux.

\begin{hypothesis}[Nonconfining charged sector]
\label{hyp:nonconfining-charged-sector-datum}
A charged sector is nonconfining, for the charged construction below, when it
has the following objects and properties.
\begin{enumerate}
\item A vacuum observable net \(\Obs\) with a positive-energy representation
and a physical charged representation \((\Hilb_q,U_q)\) of the observable net.
The representation carries a self-adjoint boundary charge \(Q\) and, for each
smooth angular test \(f\) on the sphere at spatial infinity, a Gauss-law flux
operator \(\Phi_\infty(f)\) obtained as the large-sphere limit of the
corresponding electric-flux smearing on a common invariant domain.
\item Compactly localized gauge-invariant observables preserve the boundary
charge and the angular flux sector in the large-sphere limit:
\[
  [A,Q]=0,\qquad [A,\Phi_\infty(f)]=0
\]
on the common domain whenever \(A\) is localized in a bounded spacetime
region.  Thus the charged representation is labelled by data at infinity
rather than by an operator in the bounded observable net.
\item The charged sector has finite-energy asymptotic content.  In a massive
screened or Coulomb-branch situation this means an isolated positive-energy
mass hyperboloid
\[
  \Sigma_{m,q}^+
  =
  \{p\in\mathbb R^{1,d}\mid p^2=-m^2,\ p^0>0\}
\]
inside the joint spectrum of \(U_q\), with finite-dimensional multiplicity
space \(E_{p,q}\) and spectral projection \(P_{1,q}\).  In a massless
long-range abelian theory the corresponding spectral replacement is a charged
threshold or infraparticle measure together with its flux-sector label; ordinary
isolated-shell Haag--Ruelle estimates are then replaced by the modified
asymptotic problem stated below.
\item There is a class of gauge-invariant noncompact charged creators
\(\Psi_{q,\Gamma}\), where \(\Gamma\) denotes a Wilson-line ray, Coulomb
profile, or equivalent dressing geometry reaching infinity, such that their
smeared vectors have a nonzero projection to the charged one-particle
subspace, namely
\[
  P_{1,q}\Psi_{q,\Gamma}(f)\vac\ne0
\]
in the isolated-shell case, or a nonzero threshold coefficient in the
infraparticle case.  The dressing geometry determines the angular flux label
of the vector.
\item Truncated dressings must have a finite-energy limit after the local
renormalizations belonging to the chosen charged operator.  Let
\[
  \xi_{L,\varepsilon}
  =
  \frac{\Psi_{q,\Gamma;L,\varepsilon}(f)\vac}
       {\|\Psi_{q,\Gamma;L,\varepsilon}(f)\vac\|}
\]
when the denominator is nonzero, and let \(E_q(\Delta)\) be the spectral
projection of the charged Hamiltonian \(H_q\).  The operational condition is
spectral tightness:
\begin{equation}
  \lim_{E\to\infty}
  \sup_{(L,\varepsilon)\in\mathcal A}
  \bigl\|(1-E_q([0,E]))\xi_{L,\varepsilon}\bigr\|=0
  \label{eq:charged-sector-spectral-tightness}
\end{equation}
for the admissible truncation schedules \(\mathcal A\) used in the large-time
construction.  Uniform first-moment control,
\[
  \sup_{(L,\varepsilon)\in\mathcal A}
  \langle \xi_{L,\varepsilon},H_q\xi_{L,\varepsilon}\rangle<\infty ,
\]
is a useful sufficient estimate by the spectral Markov inequality, but the
scattering construction needs the tightness of the spectral measures
themselves.  The tightness condition is what permits extraction of finite-energy
subsequential limits and makes the projection or threshold coefficient in the
previous item meaningful.

The confining alternative is a spectral lower-bound statement for the declared
class of gauge-invariant charged truncations.  Let \(\mathcal D_{q,L}\) be a
class of normalized truncated charged vectors whose boundary flux reaches
distance \(L\).  If
there are \(\sigma>0\), \(L_0\), and \(c=o(L)\) such that every
\(\xi\in\mathcal D_{q,L}\), \(L\ge L_0\), has spectral support above
\begin{equation}
  E_{\rm vac}+\sigma L-c(L),
  \label{eq:charged-sector-string-energy-lower-bound}
\end{equation}
or equivalently
\[
  \langle \xi,H_q\xi\rangle
  \ge E_{\rm vac}+\sigma L-c(L),
\]
then the family cannot satisfy
\eqref{eq:charged-sector-spectral-tightness}.  In that case the
finite-energy charged asymptotic sector is absent and the scattering problem
belongs to neutral composite sectors.
\end{enumerate}
\end{hypothesis}

The word ``nonconfining'' in the rest of this section refers precisely to
Hypothesis~\ref{hyp:nonconfining-charged-sector-datum}.  The hypothesis does
not solve the charged Haag--Ruelle problem.  It states the minimum positive
spectral and representation-theoretic data on which that problem can be
posed.  The subsequent LSZ and Dollard constructions explain what follows
from these inputs and which large-time estimates still have to be proved.

\paragraph{Truncation topology for noncompact dressings.}
The symbol \(\Psi_{q,\Gamma}\) in
Hypothesis~\ref{hyp:nonconfining-charged-sector-datum} is not an ordinary
local operator-valued distribution with a hidden bounded localization region.
For a Wilson-line ray or Coulomb dressing it must be defined as a
renormalized limit of finite dressings.  A precise charged construction must
therefore fix, before any large-time scattering limit is taken, a family
\[
  \Psi_{q,\Gamma;L,\varepsilon}(f)
  =
  N_{\Gamma;L,\varepsilon}\,
  \psi_{q,\varepsilon}(f)\,W_{\Gamma;L,\varepsilon}
\]
on a common finite-energy domain: \(L\) is the length or infrared truncation
of the dressing, \(\varepsilon\) is the ultraviolet regulator at the charged
endpoint and along cusps or corners, \(W_{\Gamma;L,\varepsilon}\) is the
finite Wilson-line or Coulombic factor, and \(N_{\Gamma;L,\varepsilon}\)
contains the endpoint, line, cusp, and local charged-field renormalizations
belonging to the chosen scheme.  The infinite dressing is a shorthand only
after matrix elements against finite-energy vectors have limits, with the
boundary charge and angular flux label held fixed:
\[
  \lim_{\varepsilon\downarrow0}\lim_{L\to\infty}
  \langle\xi,\Psi_{q,\Gamma;L,\varepsilon}(f)\eta\rangle
  =
  \langle\xi,\Psi_{q,\Gamma}(f)\eta\rangle .
\]
The order of limits may be replaced by a joint renormalization prescription,
but the prescription is part of the operator datum.

For Haag--Ruelle purposes fixed-time convergence is not enough.  The
large-time approximants sample increasingly long pieces of the Gauss-law tail,
so the theorem must specify admissible truncation schedules
\[
  L_i(t)\to\infty,\qquad \varepsilon_i(t)\downarrow0 ,
  \qquad \sigma t\to+\infty ,
\]
for every charged external leg.  Same-flux schedules and same-flux compact
dressing changes should define the same asymptotic vector only if their
modified comparison vectors obey a uniform tail condition on the scattering
domain.  In the notation used below this means that, for two admissible
same-flux choices \(L\) and \(L'\),
\begin{equation}
  \lim_{T\to\infty}\sup_{\sigma t\ge T}
  \bigl\|
    \mathcal C_{I,L}^{\sigma}(t)
    -
    \mathcal C_{I,L'}^{\sigma}(t)
  \bigr\|=0,
  \label{eq:charged-truncation-uniform-norm}
\end{equation}
and, whenever the norm derivative is used in Cook's argument,
\begin{equation}
  \lim_{T\to\infty}
  \int_{\sigma t\ge T}
  \left\|
    \frac{\dd}{\dd t}
    \bigl(
      \mathcal C_{I,L}^{\sigma}(t)
      -
      \mathcal C_{I,L'}^{\sigma}(t)
    \bigr)
  \right\|\,\dd t
  =0 .
  \label{eq:charged-truncation-uniform-cook}
\end{equation}
Equations~\eqref{eq:charged-truncation-uniform-norm} and
\eqref{eq:charged-truncation-uniform-cook} are load-bearing estimates: they
are the point where the formal infinite Wilson line becomes a legitimate
asymptotic coordinate.  A construction that only supplies
\(\Psi_{q,\Gamma;L,\varepsilon}(f)\) for each fixed \(L\) and then takes
\(L\to\infty\) at fixed time supplies a family of dressed matrix elements,
while the charged Haag--Ruelle input requires control of the joint
large-time/noncompact-tail limit.

\begin{definition}[Dressed charged LSZ problem]
\label{def:dressed-charged-lsz-problem}
A dressed charged LSZ problem consists of:
\begin{enumerate}
\item a nonconfining charged sector in the sense of
Hypothesis~\ref{hyp:nonconfining-charged-sector-datum};
\item a family of gauge-invariant noncompact charged creators
\(\Psi_{q,\gamma}(x)\), or Coulombic variants with the same asymptotic flux;
\item a specification of the dressing geometry \(\gamma\) in terms of the
asymptotic velocity or momentum labels;
\item a spectral statement for the vectors \(\Psi_{q,\gamma}(f)\vac\), either
an isolated massive shell in an infrared-regulated or screened model, or the
corresponding charged infraparticle threshold structure in massless QED;
\item a reduction prescription for Fourier-transformed dressed correlators
\[
  \langle\vac|
  T\{\Psi_{q_1,\gamma_1}(x_1)\cdots
      \Psi^\dagger_{q_n,\gamma_n}(y_n)\,\mathcal O_1\cdots\mathcal O_k\}
  |\vac\rangle
\]
which extracts the same matrix elements as the large-time dressed wave
operators.
\end{enumerate}
\end{definition}

This definition identifies the replacement for ordinary local-field LSZ in a
charged gauge sector.  The external residue, when an isolated charged shell is
available after a regulator or screening deformation, is the residue of the
dressed two-point function.  It includes the short-distance renormalization of
the matter field and the endpoint or cusp renormalization of the Wilson-line
dressing.  Different dressing geometries define different charged operators;
showing that a physical on-shell answer is independent of the unobserved
geometric choices is part of the reduction theorem, not an assumption inside
the scalar LSZ formula.

\subsection{Dressed Residues as Charged Coordinates}

Before discussing the geometry of Wilson lines, it is useful to state the
finite-dimensional algebra that any dressed LSZ theorem must implement.  Fix,
temporarily, an infrared regulator or screening deformation for which a
charged isolated massive shell exists.  Let
\[
  \{\Psi_a\}_{a=1}^r
\]
be finitely many gauge-invariant dressed charged interpolating operators with
the same charge, spin-statistics, mass, and asymptotic flux sector.  Let
\(E_p\) be the finite-dimensional one-particle multiplicity space over the
mass shell point \(p\in\Sigma_m^+\), and choose an orthonormal basis
\(\{|p,\lambda\rangle\}_{\lambda=1}^{N}\) of \(E_p\).  The one-particle
matrix coefficients of the dressed coordinates are the numbers
\[
  z_a{}^\lambda(p)
  =
  \langle p,\lambda|\Psi_a(0)\vac\rangle .
\]
Equivalently, for a positive-frequency packet \(f\) supported in the
isolating neighborhood of the shell,
\[
  P_1\Psi_a(f)\vac
  =
  \int_{\Sigma_m^+}\dd\mu_m(p)\,\widehat f(p)
  \sum_{\lambda=1}^{N}z_a{}^\lambda(p)|p,\lambda\rangle .
\]
The singular part of the dressed two-point function is therefore the
Gram matrix of these overlap vectors:
\begin{equation}
  \lim_{p^2\to -m^2}
  (p^2+m^2-\ii0)\widehat G_{ab}(p)
  =
  \ii\sum_{\lambda=1}^{N}
  z_a{}^\lambda(p)\,\overline{z_b{}^\lambda(p)}
  \equiv \ii Z_{ab}(p),
  \label{eq:dressed-lsz-residue-gram-coordinate}
\end{equation}
in the time-ordered convention used in the LSZ chapter.  Thus the pole
residue is not an intrinsic scalar attached to a raw noncompact operator; it
is the coordinate Gram matrix of the physical one-particle shell in the
chosen dressed operator coordinates.

The LSZ extraction of a normalized external charged state is possible from
this family precisely when the overlap map
\[
  z(p):E_p\longrightarrow \mathbb C^r,
  \qquad
  \eta_\lambda |p,\lambda\rangle
  \longmapsto
  \left(\sum_{\lambda}z_a{}^\lambda(p)\eta_\lambda\right)_{a=1}^r
\]
has full column rank.  A choice of left inverse
\[
  L(p):\mathbb C^r\longrightarrow E_p,
  \qquad L(p)z(p)=\mathbf 1_{E_p},
\]
then supplies the external charged residue coordinate.  If
\[
  \widehat H_a(p;\Xi)
  =
  \frac{\ii\sum_{\lambda}z_a{}^\lambda(p)\mathcal A_\lambda(p;\Xi)}
       {p^2+m^2-\ii0}
  +\widehat H_a^{\rm less}(p;\Xi)
\]
is a Fourier-transformed dressed correlator with one external charged
insertion and all other variables denoted by \(\Xi\), then
\begin{equation}
  \lim_{p^2\to -m^2}
  (p^2+m^2-\ii0)
  \sum_a L^\rho{}_a(p)\,\widehat H_a(p;\Xi)
  =
  \ii\,\mathcal A_\rho(p;\Xi).
  \label{eq:dressed-lsz-left-inverse-residue}
\end{equation}
The positive matrix \(Z^{-1/2}\) in the finite-regulator theorem below is the
special square-coordinate form of this left-inverse operation after the
operator labels have been identified with an orthonormal multiplicity basis.

This algebra also gives the exact meaning of a harmless dressing change.  If
another family is related by a finite invertible change of dressed
coordinates,
\[
  \Psi'_\alpha=\sum_a M_\alpha{}^a\Psi_a ,
\]
with the same charged shell and the same asymptotic flux sector, then
\[
  z'=Mz,\qquad Z'=MZM^\dagger,\qquad L'=LM^{-1}.
\]
Equation~\eqref{eq:dressed-lsz-left-inverse-residue} is unchanged.  This is a
coordinate statement on a fixed one-particle charged sector; it is not a
path-independence theorem for arbitrary Wilson lines.  A change of the
asymptotic ray or Coulomb tail changes the Gauss-law flux at infinity in an
unscreened theory and need not be represented by any finite invertible matrix
on a common Hilbert-space shell.  The soft-sector calculations below make this
failure explicit.

\subsection{Endpoint Renormalization and Dressed External Residues}

The overlap matrix \(z_a{}^\lambda(p)\) is not the residue of a bare local
field.  It is the residue of a renormalized nonlocal charged coordinate.  For
a Wilson-line or Coulombic coordinate, the renormalization includes the local
charged field, the charged endpoint of the dressing, any line self-energy
subtractions belonging to the chosen regulator, and cusp or corner
renormalizations whenever the dressing geometry has a nonsmooth join.  These
factors are part of the external coordinate used by dressed LSZ.

Fix one charged shell and one asymptotic flux block.  Suppose that the same
finite family of renormalized dressed coordinates is described at two
renormalization scales by an invertible matrix
\[
  R_{\mu'\mu}(p,\Gamma):
  \qquad
  \Psi_{\alpha}^{(\mu')}
  =
  \sum_a
  R_{\mu'\mu,\alpha}{}^{a}(p,\Gamma)\Psi_a^{(\mu)} .
\]
Then the one-particle overlaps and pole Gram matrix transform as
\begin{equation}
  z^{(\mu')}=R_{\mu'\mu}z^{(\mu)},
  \qquad
  Z^{(\mu')}=R_{\mu'\mu}Z^{(\mu)}R_{\mu'\mu}^{\dagger},
  \qquad
  L^{(\mu')}=L^{(\mu)}R_{\mu'\mu}^{-1}.
  \label{eq:dressed-lsz-finite-renormalization-coordinate}
\end{equation}
For an \(n\)-leg dressed correlator residue tensor this means
\begin{equation}
  \mathcal H_{a_1\cdots a_n}^{(\mu')}
  =
  (R_{1,\mu'\mu})_{a_1}{}^{b_1}\cdots
  (R_{n,\mu'\mu})_{a_n}{}^{b_n}
  \mathcal H_{b_1\cdots b_n}^{(\mu)} ,
  \label{eq:dressed-lsz-renormalized-residue-tensor}
\end{equation}
with one finite renormalization matrix for each external dressed coordinate.
Combining \eqref{eq:dressed-lsz-finite-renormalization-coordinate} and
\eqref{eq:dressed-lsz-renormalized-residue-tensor} gives
\[
  L_1^{(\mu')}\cdots L_n^{(\mu')}
  \mathcal H^{(\mu')}
  =
  L_1^{(\mu)}\cdots L_n^{(\mu)}
  \mathcal H^{(\mu)} .
\]
Thus the normalized external charged coefficient is independent of the
chosen finite endpoint/cusp renormalization only after the left inverses are
renormalized with the same nonlocal operator data as the residue tensor.

In differential form, let
\[
  \Gamma_{\ell}^{\Psi}
  =
  \mu\frac{\dd R_{\ell,\mu\mu_0}}{\dd\mu}R_{\ell,\mu\mu_0}^{-1}
\]
be the anomalous-dimension matrix of the \(\ell\)-th dressed external
coordinate within the fixed flux block.  It contains the matter-field
contribution, endpoint and cusp contributions of the Wilson-line dressing,
and possible finite-dimensional mixing among dressed coordinates with the
same charge, spin, mass, and asymptotic flux.  The residue data obey the
coordinate equations
\begin{equation}
  \mu\frac{\dd z_\ell}{\dd\mu}=\Gamma_\ell^\Psi z_\ell,
  \qquad
  \mu\frac{\dd L_\ell}{\dd\mu}=-L_\ell\Gamma_\ell^\Psi,
  \label{eq:dressed-lsz-overlap-left-inverse-rg}
\end{equation}
and
\begin{equation}
  \mu\frac{\dd}{\dd\mu}
  \mathcal H_{a_1\cdots a_n}
  =
  \sum_{\ell=1}^{n}
  (\Gamma_\ell^\Psi)_{a_\ell}{}^{b}
  \mathcal H_{a_1\cdots a_{\ell-1}b a_{\ell+1}\cdots a_n}.
  \label{eq:dressed-lsz-residue-tensor-rg}
\end{equation}
Equations
\eqref{eq:dressed-lsz-overlap-left-inverse-rg}--\eqref{eq:dressed-lsz-residue-tensor-rg}
imply
\[
  \mu\frac{\dd}{\dd\mu}
  \left(
    L_1\cdots L_n\mathcal H
  \right)
  =0 .
\]
This is the precise residue-level meaning of the statement that
\(Z_{\rm dressed}\) is different from the local matter-field \(Z_\psi\).
Using \(Z_\psi\) alone while changing the endpoint or cusp renormalization of
the Wilson-line coordinate leaves an uncanceled external factor.  That error
is not repaired by an inclusive photon sum or by a detector projection; it is
a mismatch in the charged operator whose residue is being extracted.

The same equations also delimit what is and is not a path-independence
statement.  A compact same-flux deformation of the dressing may change the
renormalized coordinates by a finite matrix of the form above, and then it is
absorbed by the left inverses.  A change of asymptotic ray, Coulomb tail, or
angular flux sector is not a finite \(R_{\mu'\mu}\) inside one block.  It
changes the asymptotic representation data and has to be compared by the
flux-sector and soft-representation analysis below.

\subsection{Wilson-Line Covariance and Boundary Charge}

The sign and orientation in
\eqref{eq:charged-wilson-line-dressed-field} are fixed by gauge covariance.
Let \(\gamma_{x,\infty}:[0,\infty)\to M\) be a regulated curve with
\(\gamma(0)=x\) and endpoint tending to spatial infinity.  With
\[
  A\mapsto A+\dd\alpha,\qquad
  \psi_q(x)\mapsto \ee^{\ii q\alpha(x)}\psi_q(x),
\]
the abelian Wilson factor
\[
  W_{q,\gamma}(x)
  =
  \exp\!\left\{\ii q\int_{\gamma_{x,\infty}} A\right\}
\]
transforms as
\[
  W_{q,\gamma}(x)
  \longmapsto
  \ee^{\ii q(\alpha(\infty)-\alpha(x))}W_{q,\gamma}(x).
\]
Therefore
\begin{equation}
  \Psi_{q,\gamma}(x)=\psi_q(x)W_{q,\gamma}(x)
  \longmapsto
  \ee^{\ii q\alpha(\infty)}\Psi_{q,\gamma}(x).
  \label{eq:charged-wilson-line-boundary-transformation}
\end{equation}
Equation~\eqref{eq:charged-wilson-line-boundary-transformation} is the
precise sense in which the operator is invariant under gauge transformations
trivial at infinity and carries charge \(q\) under the boundary gauge
transformation.  The boundary phase is not a gauge redundancy; it is the
global charge action in the charged sector.

For a nonabelian gauge field \(A=A_\mu^a t^a_R\dd x^\mu\) and a matter field
\(\psi_R(x)\) in representation \(R\), choose the parallel transporter
\[
  U_R(\infty,x;\gamma)
  =
  P\exp\!\left\{\ii\int_{\gamma_{x,\infty}} A_\mu^a t^a_R\,\dd x^\mu\right\}.
\]
Under a gauge transformation \(g\),
\[
  U_R(\infty,x;\gamma)
  \longmapsto
  g(\infty)U_R(\infty,x;\gamma)g(x)^{-1}.
\]
Thus \(U_R(\infty,x;\gamma)\psi_R(x)\) transforms only by \(g(\infty)\).  In
a sector where the boundary gauge group is treated as a global symmetry, this
is a charged operator; in a confining theory with no colored asymptotic
sectors, it is a noncompact string-like excitation rather than an external
particle creator.  This distinction is dynamical and cannot be read off from
the formal gauge covariance alone.

\paragraph{Boundary-charge selection rule for dressed correlators.}
The endpoint transformation supplies the Ward identity that fixes which
dressed correlators can enter an LSZ formula.  Consider, with a fixed
ultraviolet regulator and a fixed infrared endpoint prescription, a
time-ordered vacuum distribution
\[
  G(x_1,\ldots,x_n;\Xi)
  =
  \langle\vac|
  T\{\Psi_{q_1,\gamma_1}^{\epsilon_1}(x_1)\cdots
      \Psi_{q_n,\gamma_n}^{\epsilon_n}(x_n)\,\mathcal O(\Xi)\}
  |\vac\rangle .
\]
Here \(\epsilon_a=+1\) means \(\Psi_{q_a,\gamma_a}\),
\(\epsilon_a=-1\) means \(\Psi_{q_a,\gamma_a}^{\dagger}\), and
\(\mathcal O(\Xi)\) is a product of neutral gauge-invariant local
observables.  Under a boundary gauge transformation with constant value
\(\alpha_\infty\), the vacuum and the neutral observables are invariant while
\[
  \Psi_{q_a,\gamma_a}^{\epsilon_a}
  \longmapsto
  \exp(\ii \epsilon_a q_a\alpha_\infty)
  \Psi_{q_a,\gamma_a}^{\epsilon_a}.
\]
Therefore, as an equality of regulated distributions,
\[
  G
  =
  \exp\!\left(
    \ii\alpha_\infty\sum_{a=1}^n\epsilon_a q_a
  \right)G
  \qquad\text{for every }\alpha_\infty .
\]
If the signed boundary charge does not vanish,
\begin{equation}
  \sum_{a=1}^n\epsilon_a q_a\ne0,
  \label{eq:dressed-correlator-boundary-charge-selection}
\end{equation}
then \(G=0\).  Equivalently, a nonzero vacuum correlator of dressed charged
operators must be neutral after the charges carried by the endpoints at
infinity are included.  Matrix elements between charged superselection
sectors obey the same rule with the right side of
\eqref{eq:dressed-correlator-boundary-charge-selection} replaced by the
difference between the final and initial boundary charges.  Thus the
external charges in a dressed LSZ reduction form part of the boundary-charge
bookkeeping of the correlator itself.

For a nonabelian boundary group, the same statement is representation
theoretic.  If the endpoints of the dressed insertions transform in boundary
representations \(R_1,\ldots,R_n\), then the vacuum correlator is an invariant
tensor in
\[
  (R_1\otimes\cdots\otimes R_n)^{G_\infty}.
\]
If this invariant subspace is zero, the correlator vanishes.  If it is
nonzero, choosing a basis of invariant tensors is part of choosing the
external charged channel.  This is still separate from the dynamical question
whether the corresponding charged sector has finite-energy asymptotic
particles; in a confining phase the representation-theoretic singlet
condition may be satisfied only by neutral composite operators, and ordinary
Haag--Ruelle theory then applies to those composite particles rather than to
individual colored endpoints.

\subsection{Ray Dressings as Gauge-Slice Coordinates}

The preceding covariance statement has a second, more computational,
interpretation.  A Wilson-line dressed charged coordinate is the ordinary
charged coordinate written on a gauge slice adapted to the same ray.  The
statement below is elementary in abelian gauge theory, but it is a useful
guardrail for perturbative calculations of dressed correlators.\footnote{A
recent perturbative study of finite-distance dressed observables in
electrodynamics and perturbative gravity is C.~Cheung,
A.~Sivaramakrishnan, J.~Wilson-Gerow, and L.~Zhou, \emph{On Perturbatively
Dressed Observables}, \href{https://arxiv.org/abs/2605.26077}
{arXiv:2605.26077}.  The use made here is only the gauge-slice viewpoint; the
abelian identities and denominators in this subsection are derived directly.}

Let \(u^\mu\) be a fixed nonzero vector and let the half-line from \(x\) be
\[
  \gamma_{x,u}(\tau)=x+\tau u,\qquad \tau\ge0 .
\]
Work on a class of gauge potentials \(A\) and gauge functions \(\alpha\) for
which \(u^\mu A_\mu(x+\tau u)\) and
\(u^\mu\partial_\mu\alpha(x+\tau u)\) are absolutely integrable on
\([0,\infty)\) and for which
\[
  \alpha(x+\tau u)\longrightarrow0,
  \qquad
  u^\mu A_\mu(x+\tau u)\longrightarrow0
  \qquad(\tau\to\infty)
\]
along every ray under consideration.  Define the line functional
\begin{equation}
  \chi_u[A](x)
  =
  \int_0^\infty \dd\tau\,u^\mu A_\mu(x+\tau u).
  \label{eq:ray-dressing-line-functional}
\end{equation}
For the abelian gauge convention
\[
  A_\mu\longmapsto A_\mu+\partial_\mu\alpha,
  \qquad
  \psi_q\longmapsto \ee^{\ii q\alpha}\psi_q ,
\]
one has
\[
  \chi_u[A+\dd\alpha](x)
  =
  \chi_u[A](x)-\alpha(x).
\]
Hence
\begin{equation}
  \Psi_{q,u}(x)
  =
  \ee^{\ii q\chi_u[A](x)}\psi_q(x)
  \label{eq:ray-dressed-charged-coordinate}
\end{equation}
is invariant under gauge transformations that are trivial at the ray endpoint.

\paragraph{Ray dressing as an axial-gauge coordinate.}
For every \(A\) in the class above, set
\[
  A_\mu^{[u]}(x)=A_\mu(x)+\partial_\mu\chi_u[A](x),
  \qquad
  \psi_q^{[u]}(x)=\ee^{\ii q\chi_u[A](x)}\psi_q(x).
\]
Then
\begin{equation}
  u^\mu A_\mu^{[u]}(x)=0,
  \qquad
  \psi_q^{[u]}(x)=\Psi_{q,u}(x).
  \label{eq:ray-dressing-axial-coordinate-equality}
\end{equation}
Thus a dressed charged insertion may be computed, in any finite regularized
perturbative path integral for which this gauge slice and the corresponding
change of variables have been defined with their regulator and
Faddeev--Popov factors, as the ordinary charged field inserted in the gauge
\(u^\mu A_\mu=0\) with the same endpoint prescription.
Indeed, differentiating \eqref{eq:ray-dressing-line-functional} in the ray
direction gives
\[
\begin{aligned}
  u^\nu\partial_\nu\chi_u[A](x)
  &=
  \int_0^\infty\dd\tau\,
  \frac{d}{d\tau}\bigl(u^\mu A_\mu(x+\tau u)\bigr)
  \\
  &=
  \lim_{T\to\infty}u^\mu A_\mu(x+Tu)-u^\mu A_\mu(x)
  =
  -u^\mu A_\mu(x).
\end{aligned}
\]
Therefore
\[
  u^\mu A_\mu^{[u]}(x)
  =
  u^\mu A_\mu(x)+u^\mu\partial_\mu\chi_u[A](x)
  =
  0 .
\]
The second identity in
\eqref{eq:ray-dressing-axial-coordinate-equality} is the definition of
\(\psi_q^{[u]}\) together with \eqref{eq:ray-dressed-charged-coordinate}.
The preceding transformation law for \(\chi_u\) shows that this construction
is constant on gauge orbits once the boundary gauge value has been fixed.

The perturbative singularities produced by the ray are also fixed at this
elementary level.  Insert an endpoint regulator
\[
  \chi_{u,\epsilon}[A](x)
  =
  \int_0^\infty\dd\tau\,\ee^{-\epsilon\tau}
  u^\mu A_\mu(x+\tau u),
  \qquad \epsilon>0,
\]
and use the Fourier convention
\[
  A_\mu(y)=\int\frac{\dd^D k}{(2\pi)^D}\,
  \ee^{-\ii k\cdot y}\widehat A_\mu(k).
\]
Then
\begin{equation}
  \chi_{u,\epsilon}[A](x)
  =
  -\ii
  \int\frac{\dd^D k}{(2\pi)^D}\,
  \ee^{-\ii k\cdot x}
  \frac{u^\mu\widehat A_\mu(k)}{k\cdot u-\ii\epsilon}.
  \label{eq:finite-ray-dressing-eikonal-denominator}
\end{equation}
Indeed, substituting the Fourier transform gives
\[
  \chi_{u,\epsilon}[A](x)
  =
  \int\frac{\dd^D k}{(2\pi)^D}
  \ee^{-\ii k\cdot x}u^\mu\widehat A_\mu(k)
  \int_0^\infty\dd\tau\,
  \ee^{-(\epsilon+\ii k\cdot u)\tau},
\]
and the last integral is
\[
  \frac{1}{\epsilon+\ii k\cdot u}
  =
  -\frac{\ii}{k\cdot u-\ii\epsilon}.
\]
Expanding the exponential in
\eqref{eq:ray-dressed-charged-coordinate} therefore inserts the distribution
\((k\cdot u-\ii0)^{-1}\) for every photon emitted from the line.  These
linear denominators are part of the nonlocal composite operator and of the
corresponding axial-gauge coordinate system.  They must be treated as
additional singular hypersurfaces in loop integrals involving dressed
finite-distance correlators.  They do not, by themselves, supply the
large-time Haag--Ruelle estimate or the infrared representation theory
required for charged scattering; those are the separate problems formulated
below and in Chapter~\ref{ch:volume-ii-infrared-inclusive-qed}.

\subsection{The Missing Haag--Ruelle Estimate}

Ordinary Haag--Ruelle theory uses two facts at large time: the one-particle
part of \(B_t(h)\vac\) is time independent, and commutators of creators with
disjoint velocity supports are rapidly decreasing because the creators are
almost local.  The almost-local closure of the Gauss-law obstruction above
shows that, for an unscreened Gauss-law charge, a charged creator in the
observable net cannot satisfy that localization hypothesis.  A Wilson-line or
Coulomb-dressed charged creator has noncompact localization for structural
reasons, so the local commutator estimate must be replaced rather than reused.

The geometric replacement is not ``large spacelike separation'' of bounded
double cones.  Let \(V_h\) be the compact velocity support of the charged wave
packet and let \(\mathcal C_\Gamma\) denote the asymptotic cone, ray, or
Coulomb coordinate carried by the dressing.  A future charged theorem must
separate two different pieces of the large-time operator.  The compact core of
\(\Psi_{\Gamma,t}(h)\) is transported into the velocity tube \(tV_h+O(1)\) and
can be treated by the ordinary spacelike-separation argument when the velocity
supports are disjoint.  The Gauss-law tail attached to
\(\mathcal C_\Gamma\), however, remains noncompact and is not made harmless by
separating the cores.  Consequently the commutator estimate has to be a
decomposition into the following three separately estimated contributions:
\[
  \text{core--core spacelike term}
  \quad+\quad
  \text{tail--tail flux pairing}
  \quad+\quad
  \text{integrable remainder}.
\]
The second term is the part which becomes the Dollard or Faddeev--Kulish
comparison phase below.  Compact deformations of the dressing ray can change
the operator by neutral field-strength insertions, as in
\eqref{eq:compact-dressing-stokes}, but they do not remove the asymptotic flux
tail.  Thus the analytic estimate to be proved is a cone/flux estimate, not a
disguised application of ordinary almost locality.

The replacement has definite analytic content: a charged construction must
separate the nonintegrable long-range part of the motion from the integrable
remainder.  For a charged packet \(h\), write
\[
  \Psi_{\Gamma,t}(h)
\]
for a dressed creator with datum
\(\Gamma=(q,\gamma,\mathcal E_\infty,\mathfrak s)\): \(q\) is the boundary
charge, \(\gamma\) is the geometric dressing or Coulomb coordinate,
\(\mathcal E_\infty\) is the limiting Gauss-law flux density, and
\(\mathfrak s\) records spin and internal multiplicity labels.  A
Haag--Ruelle replacement for a finite velocity-separated list
\[
  I=((\Gamma_1,h_1),\ldots,(\Gamma_n,h_n))
\]
must provide a real locally absolutely continuous comparison phase
\(\Theta_I(t)\) and vectors
\begin{equation}
  \mathcal C_I(t)
  =
  \ee^{-\ii\Theta_I(t)}
  \Psi_{\Gamma_1,t}(h_1)\cdots\Psi_{\Gamma_n,t}(h_n)\vac .
  \label{eq:charged-hr-modified-comparison-vector}
\end{equation}
The desired estimates have the following precise content.
\begin{enumerate}
\item The one-particle or threshold coordinate of
\(\Psi_{\Gamma,t}(h)\vac\) is independent of \(t\) after the chosen
comparison dynamics is applied.  In an infrared-regulated massive charged
sector this is an ordinary isolated-shell projection.  In massless QED it is
instead a threshold or detector datum in a velocity/flux representation, not
a Wigner one-particle vector in a single photon Fock space.
\item For two velocity-separated dressed creators, the large-time exchange
relation on the finite-energy domain has the form
\begin{equation}
  \Psi_{\Gamma_i,t}^{\#}(h_i)\Psi_{\Gamma_j,t}^{\#}(h_j)
  =
  \ee^{\ii\theta_{ij}^{\#\#}(t)}
  \Psi_{\Gamma_j,t}^{\#}(h_j)\Psi_{\Gamma_i,t}^{\#}(h_i)
  +R_{ij}^{\#\#}(t),
  \qquad
  \int_{T_0}^{\infty}\|R_{ij}^{\#\#}(t)\xi\|\,\dd t<\infty ,
  \label{eq:dressed-hr-exchange-estimate}
\end{equation}
for every vector \(\xi\) in the scattering domain and for
each choice in which a superscript \(\#\) denotes either the operator itself
or its adjoint.  The phase \(\theta_{ij}^{\#\#}\), when present, is determined
by the asymptotic flux data.  In a local massive theory this phase is zero and
the remainder is rapidly decreasing in operator norm.  In a long-range
charged theory a term with \(\dot\theta_{ij}(t)\sim c_{ij}/t\) is not an
error; it must be included in the comparison dynamics.
\item The comparison phase in
\eqref{eq:charged-hr-modified-comparison-vector} cancels all nonintegrable
pair terms in the time derivative:
\begin{equation}
  \left\|
  \frac{\dd}{\dd t}\mathcal C_I(t)
  \right\|
  \le R_I(t),
  \qquad
  R_I\in L^1([T_0,\infty)).
  \label{eq:charged-hr-modified-cook-estimate}
\end{equation}
This is the charged analogue of Cook's estimate.  It is the analytic heart of
the missing theorem.
\item Scalar products of the modified vectors have limits computed in the
appropriate asymptotic representation.  For a screened theory with isolated
massive charged shells this representation may be an ordinary charged Fock
space.  For massless QED it must be a representation carrying the
velocity/flux labels discussed below; one cannot first form a single photon
Fock space and then attach the flux by a limiting unitary.
\item Dressings with the same limiting flux and a finite Hilbert-space
difference act as changes of interpolating coordinates.  Dressings with
different limiting fluxes define different charged sectors.  The distinction
is quantified below by the soft-norm and Weyl-implementation calculations.
\end{enumerate}

\paragraph{Finite-dimensional Coulomb-tail comparison.}
The following elementary calculation fixes the normalization of the
logarithmic comparison phase, but it is not a theorem about QFT.  Let
\(u\in\mathbb R^3\setminus\{0\}\), \(b\in\mathbb R^3\), \(a>0\), and
\(\kappa\in\mathbb R\), and define
\[
  V(t)=\frac{\kappa}{\sqrt{a^2+|b+ut|^2}},
  \qquad
  \Phi(T)=\int_{T_0}^{T}V(t)\,\dd t .
\]
Introduce
\[
  \sigma=\frac{b\cdot u}{|u|^2},
  \qquad
  b_\perp=b-\sigma u,
  \qquad
  \rho^2=a^2+|b_\perp|^2 .
\]
Then
\[
  a^2+|b+ut|^2=\rho^2+|u|^2(t+\sigma)^2 ,
\]
and the integral has the exact primitive
\begin{equation}
  \Phi(T)
  =
  \frac{\kappa}{|u|}
  \left[
    \operatorname{arsinh}\!\left(\frac{|u|(t+\sigma)}{\rho}\right)
  \right]_{T_0}^{T}.
  \label{eq:charged-coulomb-tail-asinh-exact}
\end{equation}
Using
\(\operatorname{arsinh} x=\log(2x)+O(x^{-2})\) as \(x\to+\infty\), one obtains
a finite constant \(C(T_0,b,u,a,\kappa)\) with
\begin{equation}
  \Phi(T)=\frac{\kappa}{|u|}\log T+C(T_0,b,u,a,\kappa)+O(T^{-1})
  \qquad (T\to+\infty).
  \label{eq:charged-coulomb-tail-log-phase}
\end{equation}
The same completed-square expression gives
\[
  V(t)=\frac{\kappa}{|u|}\frac1t+O(t^{-2}).
\]
This finite-dimensional comparison calculation does not identify the
coefficient \(\kappa\) in a QFT.  Its role is to fix the logic of the missing
charged Haag--Ruelle estimate.  If the asymptotic
flux-sector pairing of two charged packets produces an effective long-range
pair term \(V_{ij}(t)\) with leading coefficient \(\kappa_{ij}/|u_{ij}|\),
then the corresponding contribution to \(\Theta_I(t)\) must contain
\[
  \frac{\kappa_{ij}}{|u_{ij}|}\log t .
\]
Treating this logarithmic phase as an error would make the Cook derivative
nonintegrable.  The open QFT problem is precisely to derive the analogue of
\(\kappa_{ij}\), and the \(L^1\) remainder after subtracting the resulting
phase, from the noncompact Wilson-line or Coulomb-dressed charged creators.

\paragraph{Coefficient and residual estimate to be proved.}
The symbol \(\kappa_{ij}\) in the preceding comparison is not a free
normalization.  A nonperturbative charged Haag--Ruelle theorem must construct
it from the asymptotic Gauss-law tail of the chosen charged coordinates.
For two velocity-separated packet data
\((\Gamma_i,h_i)\), \((\Gamma_j,h_j)\), and for each choice of creator or
adjoint signs \(\#\#\), the theorem must produce a scalar pair kernel
\(K_{ij}^{\#\#}(t)\) and a residual operator
\(\mathcal R_{ij}^{\#\#}(t)\) on the scattering domain.  Denote by
\(\mathcal D_{ij}^{\#\#}(t)\) the pair component of the large-time exchange
defect or Cook derivative in the chosen finite-regulator construction, and
by \(\mathcal P_{ij}^{\#\#}(t)\) the corresponding ordered product with the
other dressed factors held fixed.  The required decomposition is
\begin{equation}
  \mathcal D_{ij}^{\#\#}(t)\xi
  =
  \ii\,K_{ij}^{\#\#}(t)\,
  \mathcal P_{ij}^{\#\#}(t)\xi
  +
  \mathcal R_{ij}^{\#\#}(t)\xi ,
  \label{eq:charged-pair-coefficient-residual-estimate}
\end{equation}
with
\begin{equation}
  \int_{T_0}^{\infty}
  \|\mathcal R_{ij}^{\#\#}(t)\xi\|\,\dd t<\infty
  \qquad
  \text{for every scattering-domain vector }\xi .
  \label{eq:charged-pair-residual-l1}
\end{equation}
The leading part of \(K_{ij}^{\#\#}\) is fixed by a flux-pairing functional
on the limiting tails:
\begin{equation}
  K_{ij}^{\#\#}(t)
  =
  \eta_i\eta_j
  \frac{\kappa(\Gamma_i,\Gamma_j)}
       {\sqrt{a_{ij}^2+|b_{ij}+u_{ij}t|^2}}
  +
  k_{ij}^{\#\#,\mathrm{rem}}(t),
  \qquad
  k_{ij}^{\#\#,\mathrm{rem}}\in L^1([T_0,\infty)).
  \label{eq:charged-flux-pairing-kernel}
\end{equation}
Equivalently,
\[
  K_{ij}^{\#\#}(t)
  -
  \eta_i\eta_j
  \frac{\kappa(\Gamma_i,\Gamma_j)}{|u_{ij}|\,t}
  \in L^1([T_0,\infty)).
\]
In an elementary abelian Coulomb normalization the functional
\(\kappa(\Gamma_i,\Gamma_j)\) reduces to the Coulomb constant times the
product of the two scalar charges, after the velocity and angular-flux
profiles defining \(\Gamma_i,\Gamma_j\) have been fixed.  In a more general
nonconfining gauge sector it is a matrix element, or sector label, of the
asymptotic flux pairing; it is part of the representation-theoretic charged
sector, not a parameter chosen after the scattering estimate is known.

One concrete way to state the local extraction problem is to choose a finite
angular partition \(\{\Delta_a\}\) of the large sphere and write
\[
  F_i^a=\Phi_\infty^{(i)}(\Delta_a),\qquad
  F_j^b=\Phi_\infty^{(j)}(\Delta_b).
\]
The QFT estimate must then identify, after the velocity supports and dressing
coordinates have been fixed, the finite-cell limits of a bilinear tail kernel
\(\mathsf B_{ij}^{ab}\) such that
\begin{equation}
  \kappa_{\mathcal P}(\Gamma_i,\Gamma_j)
  =
  \sum_{a,b}F_i^a\,\mathsf B_{ij}^{ab}\,F_j^b
  \longrightarrow
  \kappa(\Gamma_i,\Gamma_j)
  \label{eq:charged-angular-flux-pairing-coefficient}
\end{equation}
under refinement of the partition and admissible same-flux truncation
schedules.  The zero mode
\(\sum_a F_i^a\) is only the boundary charge.  It does not determine
\eqref{eq:charged-angular-flux-pairing-coefficient} unless the sector has
already collapsed to the elementary scalar Coulomb normalization.  Thus two
charged coordinates with the same total charge but different limiting
angular flux can have different Dollard logarithmic coefficients, and using a
charge-only coefficient would leave a genuine long-range exchange error.

The same finite-cell formulation gives the stability estimate that the
large-time theorem must prove uniformly in the regulator.  If the angular
fluxes are perturbed on the finite partition by
\(\delta F_i,\delta F_j\), then
\begin{align}
  \delta\kappa_{\mathcal P}
  &=
  \langle \delta F_i,\mathsf B_{ij}F_j\rangle
  +
  \langle F_i,\mathsf B_{ij}\delta F_j\rangle
  +
  \langle \delta F_i,\mathsf B_{ij}\delta F_j\rangle ,
  \label{eq:charged-flux-pairing-coefficient-stability}
\end{align}
with the obvious matrix notation on the finite partition.  A same-flux
coordinate change is admissible for the pair estimate only when this
coefficient error contributes an \(L^1\) derivative after the selected
large-time schedule is applied.  A nonzero limiting
\(\delta\kappa_{\mathcal P}\) is not a harmless finite renormalization: it
multiplies the pair derivative by \(1/t\) and changes the asymptotic
representation.

This distinction is load-bearing.  Replacing \(\kappa\) by
\(\kappa+\delta\kappa\) leaves an uncompensated
\(\delta\kappa/(|u_{ij}|t)\) term in the Cook derivative, and this term is
not \(L^1\).  By contrast, a compact same-flux dressing change can only add
an \(L^1\) derivative to \(K_{ij}^{\#\#}\), hence changes the comparison phase
by a finite limit and does not change the logarithmic coefficient.  Equal
velocities are excluded from this pair estimate: when \(u_{ij}=0\), the pair
must be treated as a charged cluster, threshold sector, or bound-state
coordinate rather than as two separated Dollard legs.

The many-body comparison datum has the same pairwise form.  For a
velocity-separated list \(I\), write \(v_i\) for a velocity in the support of
\(h_i\), \(u_{ij}=v_i-v_j\), and let \(\eta_i=+1\) for a charged creator and
\(\eta_i=-1\) for its adjoint.  The sign convention records the fact that
conjugating a dressed charged creator reverses its boundary charge and its
asymptotic flux.  Suppose that the finite-regulator charged dynamics, after
all short-range terms have been separated, produces pair terms of the form
\begin{equation}
  V_{ij}(t)
  =
  \eta_i\eta_j
  \frac{\kappa_{ij}}
       {\sqrt{a_{ij}^2+|b_{ij}+u_{ij}t|^2}}
  +r_{ij}(t),
  \qquad
  r_{ij}\in L^1([T_0,\infty)).
  \label{eq:charged-many-body-pair-tail-datum}
\end{equation}
Here \(a_{ij}>0\) is a regulator or transverse smearing scale,
\(b_{ij}\) is the impact-parameter coordinate in the finite comparison
model, and \(\kappa_{ij}\) is the flux-pairing coefficient that the actual
QFT estimate must derive from the Wilson-line or Coulomb-tail data.  The
corresponding comparison phase must satisfy
\begin{equation}
  \dot\Theta_I(t)
  =
  \sum_{i<j}
  \eta_i\eta_j
  \frac{\kappa_{ij}}
       {\sqrt{a_{ij}^2+|b_{ij}+u_{ij}t|^2}}
  +\dot\Theta_I^{\mathrm{fin}}(t),
  \qquad
  \dot\Theta_I^{\mathrm{fin}}\in L^1([T_0,\infty)),
  \label{eq:charged-many-body-dollard-derivative}
\end{equation}
or, equivalently at the logarithmic level,
\begin{equation}
  \Theta_I(T)
  =
  \sum_{i<j}
  \eta_i\eta_j
  \frac{\kappa_{ij}}{|u_{ij}|}\log T
  +O(1)
  \qquad (T\to+\infty).
  \label{eq:charged-many-body-dollard-log}
\end{equation}
The velocity-separation hypothesis is part of the datum: if \(u_{ij}=0\) the
logarithmic coefficient is not defined, and the pair must be treated as a
single charged cluster or as a separate infraparticle/bound-state sector.
Equations
\eqref{eq:charged-many-body-pair-tail-datum}--\eqref{eq:charged-many-body-dollard-log}
are not an independent scattering construction.  They are the finite
Dollard bookkeeping that the nonperturbative estimate must reproduce:
all \(t^{-1}\) tails must be exhausted by oriented pair fluxes, and only the
residual derivative is allowed to enter the Cook \(L^1\) remainder.

\paragraph{From pair extraction to a Cook budget.}
The preceding equations become a theorem only if the many-body derivative is
controlled after every long-range pair has been removed in the same
normalization.  A useful way to state the target is the following finite
budget.  For a velocity-separated list \(I\), suppose the derivative of the
unphased dressed product has been decomposed on the scattering domain as
\begin{equation}
  \frac{\dd}{\dd t}
  \Psi_{\Gamma_1,t}^{\sigma}(h_1)\cdots
  \Psi_{\Gamma_n,t}^{\sigma}(h_n)\vac
  =
  \ii\sum_{i<j}K_{ij}(t)\,
  \mathcal P_{ij}(t)\xi_I(t)
  +
  \mathcal R_I^{\mathrm{core}}(t),
  \label{eq:charged-many-body-pair-extracted-derivative}
\end{equation}
where \(\xi_I(t)\) denotes the ordered product vector and
\(\mathcal P_{ij}(t)\) records the bounded finite-coordinate replacement
needed to compare the \(ij\)-ordered pair with \(\xi_I(t)\).  Let
\(R_I^{\mathrm{core}}(t)\) bound
\(\|\mathcal R_I^{\mathrm{core}}(t)\|\) on the selected finite-energy
domain.  If
\[
  K_{ij}(t)
  =
  \eta_i\eta_j
  \frac{\kappa_{ij}}{\sqrt{a_{ij}^2+|b_{ij}+u_{ij}t|^2}}
  +
  k_{ij}^{\rm rem}(t)
\]
and if the comparison phase uses exactly the same leading kernels, then
\begin{equation}
  \left\|
  \frac{\dd}{\dd t}\mathcal C_I^\sigma(t)
  \right\|
  \le
  R_I^{\mathrm{core}}(t)
  +
  M_I\sum_{i<j}|k_{ij}^{\rm rem}(t)|
  +
  \sum_{i<j}B_{ij}(t),
  \label{eq:charged-many-body-cook-budget}
\end{equation}
where \(M_I\) bounds the finite pair reorderings and \(B_{ij}\) records the
small difference between the exact square-root kernel and its leading
\(|u_{ij}|^{-1}t^{-1}\) logarithmic representative, plus finite
same-flux schedule errors.  The required charged estimate is that every term
on the right is in \(L^1\), uniformly over the admissible truncation
schedules.

The formula is also a negative test.  If the phase uses
\(\kappa_{ij}+\delta\kappa_{ij}\) with
\(\delta\kappa_{ij}\to\delta\kappa\neq0\), then
\[
  \frac{\delta\kappa}{|u_{ij}|\,t}
\]
appears on the right side of
\eqref{eq:charged-many-body-cook-budget}; its dyadic integrals do not tend to
zero.  If instead \(\delta\kappa_{ij}(t)\to0\) fast enough that
\(\delta\kappa_{ij}(t)/t\in L^1\), the error is a finite same-flux coordinate
change.  Thus the theorem target is the complete matched extraction of the
logarithmic phase and its \(L^1\) remainder.
It must show, in one regulator and one angular-flux convention, that the
extracted pair kernels, the comparison phase, the dressed-correlator residue
normalization, and the detector flux sector all use the same
\(\kappa_{ij}\).  Otherwise the remaining \(1/t\) term survives as physical
long-range exchange, not as notation.

\paragraph{Scalar-product Cauchy criterion.}
The comparison phase is useful only after it produces limits of scalar
products in the charged asymptotic representation.  Suppose that two
velocity-separated lists \(I,J\) have modified comparison vectors
\(\mathcal C_I(t),\mathcal C_J(t)\) on the charged scattering domain and
that, for \(t\ge T_0\),
\[
  \|\mathcal C_I(t)\|\le M_I,\qquad
  \|\mathcal C_J(t)\|\le M_J,\qquad
  \left\|\frac{\dd}{\dd t}\mathcal C_I(t)\right\|\le R_I(t),\qquad
  \left\|\frac{\dd}{\dd t}\mathcal C_J(t)\right\|\le R_J(t),
\]
with \(R_I,R_J\in L^1([T_0,\infty))\).  Then for \(T>S\ge T_0\),
\begin{equation}
  \left|
    \langle\mathcal C_I(T),\mathcal C_J(T)\rangle
    -
    \langle\mathcal C_I(S),\mathcal C_J(S)\rangle
  \right|
  \le
  M_J\int_S^T R_I(t)\,\dd t
  +
  M_I\int_S^T R_J(t)\,\dd t .
  \label{eq:charged-scalar-product-cauchy-bound}
\end{equation}
Hence the scalar product has a limit once the modified Cook estimates are
proved on a common domain and the comparison vectors are uniformly bounded.
If the logarithmic coefficient in \(\Theta_I\) is changed by
\(\delta\neq0\), the model scalar product acquires the factor
\(\exp(\ii\delta\log t)\).  Along dyadic times this changes by
\(\exp(\ii\delta\log2)\), and the function
\(t\mapsto\exp(\ii\delta\log t)\) has no large-time limit.  Thus the
scalar-product limit is another place where the flux-pairing coefficient is
not a harmless convention: the same \(L^1\) residual that makes Cook's
estimate finite is also what makes the asymptotic inner product well
defined.

\begin{hypothesis}[Charged Dollard--Cook estimate package]
\label{hyp:charged-dollard-cook-package}
Fix a nonconfining charged sector and a dense set
\(\mathfrak D_{\mathrm{ch}}\) of charged packet data
\((\Gamma,h)\), where \(\Gamma=(q,\gamma,\mathcal E_\infty,\mathfrak s)\)
records boundary charge, dressing geometry, limiting flux, and spin/internal
labels, and where \(h\) has compact velocity support.  For every finite
velocity-separated ordered list
\[
  I=((\Gamma_1,h_1),\ldots,(\Gamma_n,h_n))
\]
and for each time direction \(\sigma\in\{+,-\}\), assume the following data
and estimates.
\begin{enumerate}
\item There are dressed approximants
\(\Psi_{\Gamma_j,t}^{\sigma}(h_j)\) acting on a common finite-energy domain
containing \(\vac\), and the product vector is defined there for all
sufficiently large \(\sigma t>0\).  These approximants are limits, in the
truncation topology described above, of finite-length and ultraviolet
regularized dressings; the estimates below are required to be stable under
admissible same-flux truncation schedules.
\item There is a real locally absolutely continuous comparison phase
\(\Theta_I^\sigma(t)\), unique up to a finite large-time limit, whose
nonintegrable part is the oriented pairwise Dollard logarithm
\eqref{eq:charged-many-body-dollard-log} whenever the long-range tail is
unscreened.
\item The modified vector
\[
  \mathcal C_I^\sigma(t)
  =
  \exp[-\ii\Theta_I^\sigma(t)]
  \Psi_{\Gamma_1,t}^{\sigma}(h_1)\cdots
  \Psi_{\Gamma_n,t}^{\sigma}(h_n)\vac
\]
is locally absolutely continuous in Hilbert norm and satisfies an integrable
Cook bound
\begin{equation}
  \left\|
    \frac{\dd}{\dd t}\mathcal C_I^\sigma(t)
  \right\|
  \le R_I^\sigma(t),
  \qquad
  R_I^\sigma\in L^1(\{t:\sigma t\ge T_0\}).
  \label{eq:charged-dollard-cook-package-integrable-derivative}
\end{equation}
\item For every pair of velocity-separated lists \(I,J\), the scalar products
have limits
\begin{equation}
  \langle J,I\rangle_{\mathrm{as}}^\sigma
  =
  \lim_{\sigma t\to+\infty}
  \langle
    \mathcal C_J^\sigma(t),
    \mathcal C_I^\sigma(t)
  \rangle ,
  \label{eq:charged-dollard-cook-asymptotic-inner-product}
\end{equation}
and these limits define a positive semidefinite sesquilinear form on the
algebraic span of charged asymptotic symbols.  The form is compatible with
finite refinements of velocity partitions and with finite same-flux dressing
coordinate changes.
\item If spacetime covariance is to be asserted, the approximants and phases
obey the corresponding covariance law modulo vectors whose norms tend to
zero in the same large-time limit.
\end{enumerate}
\end{hypothesis}

\begin{construction}[Charged wave maps from a Dollard--Cook package]
\label{cons:charged-dollard-cook-wave-maps}
Assume Hypothesis~\ref{hyp:charged-dollard-cook-package}.  Let
\(\mathcal H_{\mathrm{as},0}^{\sigma}\) be the algebraic vector space
generated by velocity-separated charged asymptotic symbols \(e_I^\sigma\),
modulo the finite linear relations imposed by packet refinement and same-flux
coordinate changes.  Equip it with the positive semidefinite form
\eqref{eq:charged-dollard-cook-asymptotic-inner-product}, quotient by its
null space, and complete.  The resulting Hilbert space is denoted
\(\mathcal H_{\mathrm{as}}^{\sigma,\mathrm{ch}}\).

For a generator \(e_I^\sigma\), define
\begin{equation}
  \Omega_{\sigma}^{\mathrm{ch}} e_I^\sigma
  =
  \lim_{\sigma t\to+\infty}\mathcal C_I^\sigma(t).
  \label{eq:charged-dollard-cook-wave-map-definition}
\end{equation}
The limit exists because, for \(\sigma t'>\sigma t\ge T_0\),
\[
  \|\mathcal C_I^\sigma(t')-\mathcal C_I^\sigma(t)\|
  \le
  \int_{\sigma t}^{\sigma t'}R_I^\sigma(\sigma s)\,\dd s
\]
with the integral taken along the chosen time ray.  The \(L^1\) tail in
\eqref{eq:charged-dollard-cook-package-integrable-derivative} tends to zero,
so \(\mathcal C_I^\sigma(t)\) is Cauchy in \(\Hilb\).

The definition is independent of null asymptotic symbols.  If
\(\xi=\sum_\alpha c_\alpha e_{I_\alpha}^\sigma\), then the corresponding
physical approximant is
\[
  \mathcal C_\xi^\sigma(t)
  =
  \sum_\alpha c_\alpha\mathcal C_{I_\alpha}^\sigma(t),
\]
and Hypothesis~\ref{hyp:charged-dollard-cook-package} gives
\begin{equation}
  \lim_{\sigma t\to+\infty}\|\mathcal C_\xi^\sigma(t)\|^2
  =
  \langle \xi,\xi\rangle_{\mathrm{as}}^\sigma .
  \label{eq:charged-dollard-cook-null-vector-check}
\end{equation}
Thus a null asymptotic vector has zero physical limit.  For two algebraic
vectors \(\xi,\eta\), the same scalar-product limit gives
\[
  \langle
    \Omega_\sigma^{\mathrm{ch}}\eta,
    \Omega_\sigma^{\mathrm{ch}}\xi
  \rangle_{\Hilb}
  =
  \langle\eta,\xi\rangle_{\mathrm{as}}^\sigma .
\]
Therefore \(\Omega_\sigma^{\mathrm{ch}}\) descends to an isometry from
\(\mathcal H_{\mathrm{as}}^{\sigma,\mathrm{ch}}\) into the physical Hilbert
space.  If the covariance clause in the hypothesis is included, the usual
intertwining relation follows by applying the same norm-limit argument to the
covariance defect.
\end{construction}

Construction~\ref{cons:charged-dollard-cook-wave-maps} is deliberately
abstract.  It proves that no additional Hilbert-space miracle is needed after
the charged estimates have been established, but it does not establish those
estimates.  The remaining QFT problem is to derive
\eqref{eq:dressed-hr-exchange-estimate} and
\eqref{eq:charged-dollard-cook-package-integrable-derivative}, together with
the scalar-product limit
\eqref{eq:charged-dollard-cook-asymptotic-inner-product}, from a
nonperturbative construction of Wilson-line or Coulomb-dressed charged
creators.  If the gauge field is screened or massive, the long-range phase
may be absent and ordinary massive scattering estimates can be recovered for
gauge-invariant composite charged operators.  For an unscreened Gauss-law
charge, the flux tail is part of the asymptotic datum, so any theorem must
include it from the start.

\paragraph{Finite null-quotient ledger.}
The quotient in Construction~\ref{cons:charged-dollard-cook-wave-maps} is not
a cosmetic step.  It is the finite-dimensional algebra which lets packet
refinements and same-flux coordinate redundancies coexist with a genuine
asymptotic Hilbert space.  Suppose three asymptotic symbols \(e_1,e_2,e_3\)
have physical large-time limits
\[
  \xi_1=(1,0),\qquad
  \xi_2=(0,1),\qquad
  \xi_3=(1,1)
\]
in a two-dimensional Hilbert space.  Their asymptotic scalar-product matrix is
\begin{equation}
  G_{ab}=\langle \xi_a,\xi_b\rangle
  =
  \begin{pmatrix}
    1&0&1\\
    0&1&1\\
    1&1&2
  \end{pmatrix}.
  \label{eq:charged-asymptotic-gram-finite-cell}
\end{equation}
The matrix is positive semidefinite, not positive definite.  Its null vector
is
\begin{equation}
  n=e_1+e_2-e_3,
  \qquad
  G n=0 .
  \label{eq:charged-asymptotic-null-relation-cell}
\end{equation}
Thus \(e_3\) is the same asymptotic vector as the refinement \(e_1+e_2\).
For any coefficient vector \(c=(c_1,c_2,c_3)\),
\[
  c^*Gc
  =
  |c_1+c_3|^2+|c_2+c_3|^2
  =
  \left\|
    c_1\xi_1+c_2\xi_2+c_3\xi_3
  \right\|^2 .
\]
Quotienting by the null relation is therefore exactly what makes the
wave-map assignment \(e_a\mapsto\xi_a\) well defined and isometric.  If
\((c_1,c_2,c_3)\) is replaced by
\((c_1+c_3,c_2+c_3,0)\), the two coefficient vectors differ by a multiple of
\(n\) and define the same asymptotic state.  In the charged theorem this
finite cell is replaced by the directed family of packet refinements and
same-flux coordinate relations.  The nontrivial analytic work is still the
existence of the limits \(G_{ab}\) from Wilson-line or Coulomb-dressed
creators; once those limits are supplied, the null quotient is ordinary
Hilbert-space bookkeeping.

\begin{proposition}[Same-flux schedule invariance of charged wave maps]
\label{prop:charged-same-flux-schedule-invariance}
Assume Hypothesis~\ref{hyp:charged-dollard-cook-package}.  Let
\(\mathcal S\) and \(\mathcal S'\) be two admissible truncation schedules for
the same charged packet data, with the same boundary charges and the same
limiting angular fluxes.  Write
\(\mathcal C_{I,\mathcal S}^{\sigma}(t)\) and
\(\mathcal C_{I,\mathcal S'}^{\sigma}(t)\) for the corresponding modified
comparison vectors.  Suppose that, for every generator \(e_I^\sigma\), there is
a real locally absolutely continuous function \(\alpha_I^\sigma(t)\) with a
finite limit \(\alpha_I^\sigma(\infty)\) and
\begin{equation}
  d_I^\sigma(T)
  :=
  \sup_{\sigma t\ge T}
  \left\|
    \mathcal C_{I,\mathcal S}^{\sigma}(t)
    -
    \ee^{-\ii\alpha_I^\sigma(t)}
    \mathcal C_{I,\mathcal S'}^{\sigma}(t)
  \right\|
  \longrightarrow 0
  \qquad (T\to\infty).
  \label{eq:charged-same-flux-schedule-tail}
\end{equation}
Then the two schedules define the same physical charged wave map after the
finite phase coordinate identification
\(e_{I,\mathcal S}^{\sigma}\sim
\ee^{-\ii\alpha_I^\sigma(\infty)}e_{I,\mathcal S'}^{\sigma}\):
\begin{equation}
  \Omega_{\sigma,\mathcal S}^{\mathrm{ch}}e_I^\sigma
  =
  \ee^{-\ii\alpha_I^\sigma(\infty)}
  \Omega_{\sigma,\mathcal S'}^{\mathrm{ch}}e_I^\sigma .
  \label{eq:charged-same-flux-wave-map-invariance}
\end{equation}
For two generators \(e_I^\sigma,e_J^\sigma\), the asymptotic Gram forms obey
\begin{equation}
  \langle J,I\rangle_{\mathrm{as},\mathcal S}^{\sigma}
  =
  \ee^{\ii\alpha_J^\sigma(\infty)}
  \ee^{-\ii\alpha_I^\sigma(\infty)}
  \langle J,I\rangle_{\mathrm{as},\mathcal S'}^{\sigma}.
  \label{eq:charged-same-flux-gram-invariance}
\end{equation}
Consequently same-flux schedule changes are quotient data, not new charged
asymptotic sectors.

If the dressed-correlator interface of
Hypothesis~\ref{hyp:dressed-correlator-wave-map-interface} is also available
for the two schedules, and the external overlap maps on each finite coordinate
frame are related by
\[
  z_{\ell,\mathcal S'}(p;t)
  =
  M_\ell(t) z_{\ell,\mathcal S}(p;t)+o(1),
  \qquad
  M_\ell(t)\longrightarrow M_\ell
\]
with \(M_\ell(t)\) invertible for all sufficiently large \(t\) and
\(M_\ell\) invertible on the chosen same-flux coordinate frame, then the
left-inverse residue extraction gives the same coefficient
\(\mathcal A_{\boldsymbol\lambda}\) in the simultaneous external-shell limit.
\end{proposition}

\begin{proof}
The Cook bounds in
\eqref{eq:charged-dollard-cook-package-integrable-derivative} give Hilbert
limits for both schedules and uniform large-time bounds for the comparison
vectors.  For \(\sigma t\) large,
\[
\begin{aligned}
&
\left\|
  \Omega_{\sigma,\mathcal S}^{\mathrm{ch}}e_I^\sigma
  -
  \ee^{-\ii\alpha_I^\sigma(\infty)}
  \Omega_{\sigma,\mathcal S'}^{\mathrm{ch}}e_I^\sigma
\right\|
\\
&\quad\le
\left\|
  \Omega_{\sigma,\mathcal S}^{\mathrm{ch}}e_I^\sigma
  -
  \mathcal C_{I,\mathcal S}^{\sigma}(t)
\right\|
+
d_I^\sigma(\sigma t)
+
\left|
  \ee^{-\ii\alpha_I^\sigma(t)}
  -
  \ee^{-\ii\alpha_I^\sigma(\infty)}
\right|
\,
\left\|\mathcal C_{I,\mathcal S'}^{\sigma}(t)\right\|
\\
&\qquad+
\left\|
  \mathcal C_{I,\mathcal S'}^{\sigma}(t)
  -
  \Omega_{\sigma,\mathcal S'}^{\mathrm{ch}}e_I^\sigma
\right\| .
\end{aligned}
\]
Every term on the right tends to zero.  This proves
\eqref{eq:charged-same-flux-wave-map-invariance}.

The Gram statement is the same estimate before the last limit is taken.  Put
\[
  \widetilde{\mathcal C}_{I,\mathcal S'}^\sigma(t)
  =
  \ee^{-\ii\alpha_I^\sigma(t)}
  \mathcal C_{I,\mathcal S'}^\sigma(t).
\]
If \(M_I,M_J\) are large-time norm bounds for the corresponding comparison
vectors, then
\[
\begin{aligned}
&
\left|
  \langle
    \mathcal C_{J,\mathcal S}^{\sigma}(t),
    \mathcal C_{I,\mathcal S}^{\sigma}(t)
  \rangle
  -
  \langle
    \widetilde{\mathcal C}_{J,\mathcal S'}^\sigma(t),
    \widetilde{\mathcal C}_{I,\mathcal S'}^\sigma(t)
  \rangle
\right|
\\
&\qquad\le
M_I d_J^\sigma(\sigma t)+M_J d_I^\sigma(\sigma t).
\end{aligned}
\]
Taking the large-time limit and then replacing
\(\alpha_I^\sigma(t),\alpha_J^\sigma(t)\) by their finite limits gives
\eqref{eq:charged-same-flux-gram-invariance}.  The null quotient therefore
identifies the two schedules by a finite coordinate phase.

For the residue statement, at finite regulator the singular tensor in
\eqref{eq:dressed-correlator-external-singular-expansion} transforms on the
\(\ell\)-th external coordinate by \(M_\ell(t)\), while the left inverse
transforms by \(L_{\ell,\mathcal S}(p;t)M_\ell(t)^{-1}\), for all sufficiently
large \(t\).  Contracting the external coordinates gives
\[
  \bigotimes_{\ell}
  \bigl(L_{\ell,\mathcal S}M_\ell(t)^{-1}\bigr)
  \,
  \bigotimes_{\ell}M_\ell(t)
  =
  \bigotimes_{\ell}L_{\ell,\mathcal S}.
\]
The \(o(1)\) schedule errors are less singular in the declared shell topology
and are killed by the same left-inverse contraction in the simultaneous
external-shell limit.  Hence the extracted coefficient is unchanged.
\end{proof}

\subsection{Dressed Correlators and Wave-Map Matrix Elements}

The wave-map construction above is formulated in Hilbert space.  A
Wilson-line LSZ theorem must also explain how it is read from
time-ordered dressed correlators.  This is a separate input: a pole in a
formal gauge-fixed charged-field propagator is not a gauge-invariant
observable, and a Wilson-line correlator is not automatically controlled by
the ordinary local-field integration-by-parts proof.  The object to be
proved is the following interface between dressed correlators and the
charged wave maps.

\begin{hypothesis}[Dressed-correlator reduction interface]
\label{hyp:dressed-correlator-wave-map-interface}
Fix a nonconfining charged sector and assume the charged
Dollard--Cook package of Hypothesis~\ref{hyp:charged-dollard-cook-package}
for the packet data under consideration.  For each external charged leg
\(\ell\), choose a finite family of renormalized finite-truncation dressed
coordinates
\[
  \Psi^{(\ell)}_{a_\ell;L_\ell,\varepsilon_\ell}(x),
  \qquad a_\ell=1,\ldots,r_\ell,
\]
converging in the truncation topology described above to a noncompact
dressing with fixed charge, asymptotic flux, and velocity support.  In an
infrared-regulated or screened setting with an isolated charged shell, let
\(E_\ell(p_\ell)\) be the corresponding one-particle multiplicity space,
let
\[
  z_\ell(p_\ell):E_\ell(p_\ell)\longrightarrow\mathbb C^{r_\ell}
\]
be the dressed one-particle overlap map, and choose a left inverse
\(L_\ell(p_\ell)z_\ell(p_\ell)=\mathbf 1_{E_\ell(p_\ell)}\).  The
dressed-correlator reduction interface consists of the following assertions,
with estimates uniform over admissible same-flux truncation schedules.
\begin{enumerate}
\item The renormalized time-ordered correlators of the finite-truncation
dressed coordinates have boundary values as tempered distributions in the
external momenta after wave-packet smearing and after the declared endpoint,
line, cusp, and local charged-field counterterms have been included.
\item Near the product of the isolated charged mass shells, their simultaneous
external singular part has finite rank in the dressed coordinates:
\begin{equation}
  \widehat G_{\mathbf a}(\mathbf p;\Xi)
  =
  \frac{\ii^{N}
    \sum_{\boldsymbol\lambda}
      \prod_{\ell=1}^{N}
        z_{\ell,a_\ell}{}^{\lambda_\ell}(p_\ell)\,
      \mathcal A_{\boldsymbol\lambda}(\mathbf p;\Xi)}
       {\prod_{\ell=1}^{N}(p_\ell^2+m_\ell^2-\ii0)}
  +
  \widehat G_{\mathbf a}^{\rm less}(\mathbf p;\Xi).
  \label{eq:dressed-correlator-external-singular-expansion}
\end{equation}
Here \(\Xi\) denotes compact gauge-invariant insertions or the remaining
external variables, and ``less'' means that after multiplying by
\(\prod_{\ell}(p_\ell^2+m_\ell^2-\ii0)\) and applying the left inverses, the
term has zero simultaneous shell limit in the wave-packet topology being
used.  Partial shell singularities are assigned to the corresponding
lower-particle disconnected or lower-channel matrix elements before the
connected amplitude is read off.
\item The coefficient
\(\mathcal A_{\boldsymbol\lambda}\) obtained from
\eqref{eq:dressed-correlator-external-singular-expansion} is the charged
wave-map matrix element:
\begin{equation}
  \mathcal A_{\boldsymbol\lambda}(\mathbf p;\Xi)
  =
  \bigl\langle
    \Omega_{+}^{\mathrm{ch}}e_{\mathrm{out},\boldsymbol\lambda},
    \mathcal O_\Xi\,
    \Omega_{-}^{\mathrm{ch}}e_{\mathrm{in},\boldsymbol\lambda}
  \bigr\rangle_{\rm conn},
  \label{eq:dressed-correlator-wave-map-matrix-element}
\end{equation}
with the total momentum delta distribution and the standard disconnected
vacuum factors removed.  The Dollard phases appearing in
Hypothesis~\ref{hyp:charged-dollard-cook-package} are part of the definition
of the asymptotic symbols \(e_{\mathrm{in/out},\boldsymbol\lambda}\); without
them a long-range charged correlator need not have a simple isolated-pole
LSZ limit.
\item Endpoint and collision contact terms created by differentiating
time-ordered Wilson-line insertions are either absorbed into the declared
renormalized dressed coordinates, assigned to lower-channel contact
coordinates, or shown to be less singular in the simultaneous external-shell
sense above.  They are not allowed to contribute an additional independent
charged one-particle residue.
\item The extraction is covariant under finite same-flux coordinate changes:
if \(z_\ell\) is replaced by \(M_\ell z_\ell\), with \(M_\ell\) invertible on
the chosen coordinate frame, then \(L_\ell\) is replaced by
\(L_\ell M_\ell^{-1}\), the residue tensor is multiplied by \(M_\ell\) on
that external coordinate, and the coefficient
\(\mathcal A_{\boldsymbol\lambda}\) is unchanged.
\end{enumerate}
\end{hypothesis}

This hypothesis is not a theorem about gauge theory.  It names the analytic
work that a Wilson-line LSZ theorem must do.  One must construct the
boundary-value distributions for the noncompact dressed insertions, prove
uniformity in the truncation schedules \(L_\ell(t)\) and
\(\varepsilon_\ell(t)\), identify the finite-rank external singular part,
control contact terms from the endpoints and cusps of the lines, and match
the resulting coefficient to the Dollard--Cook wave maps.  In massless QED
the isolated pole in
\eqref{eq:dressed-correlator-external-singular-expansion} is expected to be
replaced by a threshold coefficient in a velocity/flux representation; the
same interface then has to be rewritten with the threshold functional in
place of the mass-shell residue.

\begin{controlledapproximation}[Soft-subtracted boundary-value extraction]
\label{ca:dressed-boundary-value-hard-coefficient}
In a screened theory, or at a positive infrared cutoff before the detector
limit is taken, the correlator interface can be tested by a shell-window
extraction which separates the hard charged coefficient from soft and
endpoint data.  Let
\(\Delta_\ell(p_\ell)=p_\ell^2+m_\ell^2-\ii0\), choose wave packets
\(f_\delta\) supported in a simultaneous external-shell window of size
\(\delta\), and define
\begin{equation}
  \widehat{\mathcal B}_{\mathbf L,\varepsilon,\lambda,T}(\delta)
  =
  \left\langle
    f_\delta,\,
    \left(\bigotimes_{\ell=1}^{N}L_\ell(p_\ell)\right)
    \prod_{\ell=1}^{N}\Delta_\ell(p_\ell)\,
    \widehat G_{\mathbf a;\mathbf L,\varepsilon,\lambda,T}^{\rm ren}
  \right\rangle .
  \label{eq:dressed-boundary-value-shell-functional}
\end{equation}
Here the renormalized finite-truncation correlator includes the endpoint,
line, cusp, and local charged-field counterterms already declared in
Hypothesis~\ref{hyp:dressed-correlator-wave-map-interface}, and \(T\) records
the large-time comparison used to remove the Dollard phase.  A boundary-value
extraction of the hard flux-resolved coefficient has the form
\begin{equation}
  \widehat{\mathcal B}_{\mathbf L,\varepsilon,\lambda,T}(\delta)
  =
  \mathcal S_{\lambda,E_T}^{\rm amp}\,
  \ee^{\ii\Phi_{\Theta}(T)}
  \mathcal M_{\beta\alpha}^{\mathfrak F',\mathfrak F}
  +
  r_{\rm sh}(\delta)
  +
  r_{\rm ct}(\varepsilon,L)
  +
  r_{\rm th}(\lambda,E_T,\delta)
  +
  r_{\rm sch}(L,T).
  \label{eq:dressed-boundary-value-hard-plus-residuals}
\end{equation}
The amplitude soft factor
\(\mathcal S_{\lambda,E_T}^{\rm amp}\) is the square root, with its finite
matching phase, of the leading detector soft factor used below; at leading
homogeneous order
\[
  \left|\mathcal S_{\lambda,E_T}^{\rm amp}\right|^2
  =
  \exp\{-\|P_{E_T,M}F_{\beta\alpha}\|^2\}.
\]
The phase \(\Phi_{\Theta}(T)\) is the phase determined by the same
pair coefficients \(\kappa_{ij}\) that enter the charged Dollard--Cook
package.  The residuals have distinct meanings: \(r_{\rm sh}\) is killed by
the simultaneous shell limit, \(r_{\rm ct}\) is the remaining endpoint/cusp or
collision contact contribution after the declared counterterms and
lower-channel coordinates have been removed, \(r_{\rm th}\) is the finite
soft-threshold remainder left by the cutoff and detector window, and
\(r_{\rm sch}\) is the same-flux truncation-schedule error.

If the estimates
\[
  \begin{aligned}
    |r_{\rm sh}(\delta)|
    &\le C_{\rm sh}\delta,
    &
    |r_{\rm ct}(\varepsilon,L)|
    &\le C_{\rm ct}\rho_{\rm ct}(\varepsilon,L),
    \\
    |r_{\rm th}(\lambda,E_T,\delta)|
    &\le C_{\rm th}\rho_{\rm th},
    &
    |r_{\rm sch}(L,T)|
    &\le C_{\rm sch}\rho_{\rm sch}(L,T)
  \end{aligned}
\]
hold in the chosen wave-packet topology, then
\begin{equation}
  \widehat{\mathcal M}_{\beta\alpha}^{\mathfrak F',\mathfrak F}
  =
  \left(\mathcal S_{\lambda,E_T}^{\rm amp}\right)^{-1}
  \ee^{-\ii\Phi_\Theta(T)}
  \widehat{\mathcal B}_{\mathbf L,\varepsilon,\lambda,T}(\delta)
  \label{eq:dressed-boundary-value-hard-estimator}
\end{equation}
satisfies
\begin{equation}
  \left|
    \widehat{\mathcal M}_{\beta\alpha}^{\mathfrak F',\mathfrak F}
    -
    \mathcal M_{\beta\alpha}^{\mathfrak F',\mathfrak F}
  \right|
  \le
  \left|\mathcal S_{\lambda,E_T}^{\rm amp}\right|^{-1}
  \left(
    C_{\rm sh}\delta
    +
    C_{\rm ct}\rho_{\rm ct}
    +
    C_{\rm th}\rho_{\rm th}
    +
    C_{\rm sch}\rho_{\rm sch}
  \right).
  \label{eq:dressed-boundary-value-hard-estimator-bound}
\end{equation}
This is a fixed-soft-window estimate unless an additional relative
uniformity statement is supplied.  At fixed detector resolution and infrared
regulator, with
\(|\mathcal S_{\lambda,E_T}^{\rm amp}|\ge s_*>0\), one first removes the
long-range comparison phase and finite soft factor, then takes the shell,
endpoint/contact, threshold-window, and schedule limits in a way that makes
the numerator on the right of
\eqref{eq:dressed-boundary-value-hard-estimator-bound} vanish.  Only after
that hard coefficient has been extracted at fixed resolution may one compare
the family as \(E_T\downarrow0\).  If a simultaneous detector-threshold limit
\((\lambda_\eta,E_{T,\eta},\delta_\eta,\varepsilon_\eta,L_\eta,T_\eta)\) is
intended, the required hypothesis is instead the relative estimate
\[
  C_{\rm sh}\delta_\eta
  +C_{\rm ct}\rho_{\rm ct,\eta}
  +C_{\rm th}\rho_{\rm th,\eta}
  +C_{\rm sch}\rho_{\rm sch,\eta}
  \le
  \omega(\eta)\,
  \left|\mathcal S_{\lambda_\eta,E_{T,\eta}}^{\rm amp}\right|,
  \qquad \omega(\eta)\to0.
\]
Absolute smallness of the pre-divided residual is not enough when
\(|\mathcal S_{\lambda,E_T}^{\rm amp}|\) itself tends to zero.  Thus the pole
residue by itself is not yet the charged scattering amplitude; the soft
factor division is a conditioned extraction step, not a harmless
normalization after all infrared limits have already been taken.

This controlled approximation also gives the relevant negative tests.  If
\(\mathcal S_{\lambda,E_T}^{\rm amp}\) is not divided out, the extracted
quantity is the no-resolved-soft-photon boundary value, not the hard charged
coefficient.  If the Dollard phase uses a coefficient
\(\kappa_{ij}+\delta\kappa_{ij}\) with a nonzero limiting error, the
remaining factor \(\exp(\ii\delta\kappa_{ij}\log T)\) has no boundary-value
limit along large dyadic times.  If an endpoint or collision contact term has
the same simultaneous simple-pole rank as the external residue and is neither
absorbed into the renormalized dressed coordinate nor assigned to a
lower-channel coordinate, it shifts
\(\mathcal M_{\beta\alpha}^{\mathfrak F',\mathfrak F}\).  If
\(|\mathcal S_{\lambda,E_T}^{\rm amp}|\to0\) while the pre-divided residual is
only \(O(|\mathcal S_{\lambda,E_T}^{\rm amp}|)\), the extracted hard
coefficient retains an \(O(1)\) ambiguity even though the boundary-value
residual tends to zero.  Finally, a
soft-threshold branch term which is not less singular in the shell topology
means the isolated-shell formula has failed; in massless QED the target must
then be the flux-resolved threshold functional rather than
\eqref{eq:dressed-boundary-value-hard-estimator}.
\end{controlledapproximation}

\subsection{Finite-Regulator Dressed LSZ}

There is nevertheless a precise finite-regulator statement.  It is not the
full massless charged scattering theorem; it is the part of the desired theory
that follows from the usual LSZ argument once a dressed charged creator and an
isolated charged shell have been supplied.  Equivalently, it is the
isolated-shell realization of
Hypothesis~\ref{hyp:dressed-correlator-wave-map-interface} under the stronger
regularity assumptions listed below.

\begin{theorem}[Finite-regulator dressed LSZ]
\label{thm:finite-regulator-dressed-charged-lsz}
Fix a gauge theory with a regulator or screening deformation for which:
\begin{enumerate}
\item the physical Hilbert space contains an isolated massive charged
one-particle shell \(\Sigma_m^+\) in a charged sector;
\item for each external charged sector and mass there is a finite family of
translation-covariant gauge-invariant dressed operators \(\Psi_a(x)\) of the
form \eqref{eq:charged-wilson-line-dressed-field}, or Coulombic variants,
whose smearings are defined on a common dense domain and whose one-particle
overlap map has full rank on the chosen charged one-particle multiplicity
space;
\item the dressed two-point distribution has, in a neighborhood of
\(\Sigma_m^+\), the pole decomposition
\begin{equation}
  \widehat G_{ab}(p)
  =
  \int \dd^D x\,\ee^{\ii p\cdot x}
  \langle\vac|T\{\Psi_a(x)\Psi_b^\dagger(0)\}|\vac\rangle
  =
  \frac{\ii Z_{ab}}{p^2+m^2-\ii0}
  +\widehat G^{\rm reg}_{ab}(p),
  \label{eq:dressed-two-point-pole}
\end{equation}
where \(Z=(Z_{ab})\) is the positive Gram matrix of those one-particle
overlaps and \(\widehat G^{\rm reg}\) is less singular at the shell;
\item the large-time wave operators built from these dressed creators exist,
and time-ordered dressed correlators have the usual boundary-value and
wave-packet regularity needed for the LSZ integration by parts.
\end{enumerate}
For the displayed square-root formula choose, for each external charged
one-particle multiplicity space, a square subfamily of dressed coordinates
whose overlap matrix is invertible; equivalently restrict \(Z\) to the range
spanned by that frame and regard it as a positive definite matrix there.  If
one keeps a rectangular overcomplete family, replace every \(Z^{-1/2}\) below
by a chosen left inverse of the overlap map as in
\eqref{eq:dressed-lsz-left-inverse-residue}.  With the square-frame
convention, let \(Z^{-1/2}\) be the positive inverse square root of this
positive definite Gram matrix.  Then the simultaneous external residue of the
dressed time-ordered \((m+n)\)-point function equals the matrix element of the
scattering operator between the dressed asymptotic charged states:
\begin{equation}
\begin{aligned}
&\prod_{r=1}^m
  \bigl[(Z^{-1/2})_{a'_r b'_r}(p'_r{}^2+m_r^2)\bigr]
 \prod_{s=1}^n
  \bigl[(p_s^2+m_s^2)(Z^{-1/2})_{b_s a_s}\bigr]
  \,\widehat G^{\,b'_1\cdots b'_m}{}_{b_1\cdots b_n}
     (p'_1,\ldots;p_1,\ldots)
\\
&\qquad\longrightarrow
  \langle
    p'_1,a'_1;\ldots;p'_m,a'_m,{\rm out}
  |S|
    p_1,a_1;\ldots;p_n,a_n,{\rm in}
  \rangle_{\rm conn},
  \label{eq:finite-regulator-dressed-lsz-formula}
\end{aligned}
\end{equation}
after wave-packet smearing and removal of the total momentum-conservation
delta distribution, with the same connected/disconnected convention as in the
ordinary LSZ chapter.
\end{theorem}

\begin{proof}
The hypotheses put the dressed fields in exactly the analytic position in
which LSZ is a residue calculation on the one-particle spectral summand.  We
spell out the reduction because the fields \(\Psi_a\) are not assumed to be
localized in bounded spacetime regions.  Choose a positive-frequency
Klein--Gordon wave packet \(f\) whose Fourier transform is supported in the
isolating neighborhood of the shell \(p^2+m^2=0\).  The pole decomposition
\eqref{eq:dressed-two-point-pole} says that the distribution
\(\langle\vac|\Psi_b^\dagger(g)\Psi_a(f)|\vac\rangle\) has one-particle
part
\[
  \sum_{c,d}
  \langle g,c|f,d\rangle\, Z_{ad}\,\delta_{cb},
\]
in the mass-shell representation, with the regular part contributing a smooth
function of the external invariant.  Equivalently,
\begin{equation}
  P_1\Psi_a(f)\vac
  =
  \sum_b (Z^{1/2})_a{}^b\,|f,b\rangle ,
  \label{eq:dressed-one-particle-projection-in-proof}
\end{equation}
after choosing the positive square root on the finite multiplicity space.

For one incoming external leg, multiplication of the Fourier transform by
\((p^2+m^2)\) is the Fourier transform of
\((-\Box+m^2)\) in the corresponding coordinate, in the present mostly-plus
Lorentzian convention.  Pair this differentiated coordinate with the wave
packet \(f\) and integrate over the time slab \([-T,T]\).  On every open time
ordering chamber the differentiated time-ordered correlator is the
Klein--Gordon operator applied to the corresponding distributional matrix
element.  The spectral regularity assumption implies that the part orthogonal
to the isolated one-particle summand remains regular at the shell; after the
external factor \((p^2+m^2)\) is taken to the shell, that regular part gives
zero.  The integrations by parts therefore leave only the slab boundaries
and the standard equal-time contact terms.  The latter are local in the
external time variable and are again regular functions of the external
invariant after wave-packet smearing, so they disappear in the simultaneous
mass-shell residue.

The boundary term at \(t=-T\) is, by the assumed existence of the dressed
Haag--Ruelle limit, the incoming wave operator applied to the one-particle
vector \eqref{eq:dressed-one-particle-projection-in-proof}; the boundary term
at \(t=+T\) gives the outgoing wave operator.  Repeating the argument for all
external legs, with the usual induction over time orderings and with
disconnected vacuum components subtracted after the total momentum delta
distribution is removed, gives the scattering matrix element written in the
dressed one-particle coordinates.  Multiplication by \(Z^{-1/2}\) on every
external multiplicity index converts those coordinates to the orthonormal
mass-shell basis, and yields
\eqref{eq:finite-regulator-dressed-lsz-formula}.  No step requires
\(\Psi_a\) to be a bounded-region local observable; the required substitute
for locality is exactly the assumed existence and regularity of the dressed
large-time limits.
\end{proof}

The theorem is useful because it isolates what is known from what remains
open.  At finite infrared regulator or in a screened phase, the formal
dressed-correlator residue has a precise Hilbert-space meaning if the listed
hypotheses hold.  Removing the regulator in four-dimensional QED is a
different operation: the pole in \eqref{eq:dressed-two-point-pole} is expected
to be replaced by charged threshold behavior, and the asymptotic Hilbert space
must remember the angular electric-flux sector.

\subsection{Dressing Dependence as Coordinate Dependence}

Path dependence has two different meanings.  A deformation that changes the
field at infinity changes the charged sector.  A compact deformation, or a
replacement of one short-distance dressing by another with the same
asymptotic flux, changes the interpolating coordinate used to reach the same
charged state.

\paragraph{Compact dressing changes as LSZ-coordinate changes.}
In the finite-regulator setting of
Theorem~\ref{thm:finite-regulator-dressed-charged-lsz}, let
\(\Psi_a\) and \(\Psi'_a\) be two families of dressed charged interpolating
operators with the same charge, spin, mass, and asymptotic flux.  Suppose
their one-particle matrix elements are related by an invertible matrix \(M\),
\[
  P_1\Psi'_a(f)\vac
  =
  M_a{}^b\,P_1\Psi_b(f)\vac
  +\text{terms orthogonal to }\Hilb_1
\]
for all wave packets \(f\) supported near the shell.  Then the LSZ residues
computed with \(\Psi'\) give the same scattering matrix after the external
normalization matrix \(Z'^{-1/2}\) is used.
The singular part of the two-point function transforms as
\[
  Z' = M Z M^\dagger .
\]
The amputated external coordinates obtained from \(\Psi'\) are therefore
related to those obtained from \(\Psi\) by the unitary polar part of
\(MZ^{1/2}\) acting inside the one-particle multiplicity space.  LSZ computes
matrix elements between orthonormal one-particle vectors, not between the raw
operator labels.  After multiplying by \(Z'^{-1/2}\), this coordinate change
is removed on every external leg.  Contributions orthogonal to \(\Hilb_1\)
are regular at the isolated shell and vanish in the external residue.

For abelian Wilson lines this statement has a concrete finite-regulator
meaning.  Let \(\gamma\) and \(\gamma'\) be two oriented paths beginning at
\(x\) and sharing the same asymptotic ray.  Truncate both paths at a large
radius and use the same endpoint cap at the truncation surface.  If the
difference of the truncated paths is the boundary of an oriented compact
surface \(S_{\gamma',\gamma}\),
\[
  \partial S_{\gamma',\gamma}
  =
  \gamma'_R-\gamma_R ,
\]
then, for a smooth regulated connection \(A\) with curvature \(F=\dd A\),
Stokes' identity gives
\begin{equation}
  \int_{\gamma'_R} A-\int_{\gamma_R}A
  =
  \int_{S_{\gamma',\gamma}} F .
  \label{eq:compact-dressing-stokes}
\end{equation}
Hence
\[
  \Psi_{q,\gamma'}(x)
  =
  \Psi_{q,\gamma}(x)\,
  \exp\left\{\ii q\int_{S_{\gamma',\gamma}}F\right\}
\]
in the abelian regulated theory, with the same short-distance prescription at
the common starting point.  The extra factor is gauge invariant, neutral, and
localized in the compact region swept out by the deformation.  It can change
the overlap vector \(P_1\Psi_{q,\gamma}(f)\vac\) by a finite amount, but it
does not change the Gauss-law boundary charge or the asymptotic flux label.
This is the precise setting in which the preceding LSZ coordinate calculation
applies.

For a nonabelian connection the corresponding infinitesimal statement is a
curvature-insertion formula rather than an ordinary scalar Stokes identity.
With \(U_\gamma(s_2,s_1)\) defined by
\[
  \partial_{s_2}U_\gamma(s_2,s_1)
  =
  -\dot\gamma^\mu(s_2)A_\mu(\gamma(s_2))U_\gamma(s_2,s_1),
  \qquad U_\gamma(s_1,s_1)=1,
\]
and with a compact path variation \(\eta^\nu(s)\) that vanishes at the
endpoint and on the asymptotic tail, one has
\[
  \delta U_\gamma(\infty,0)
  =
  \int_0^\infty\dd s\,
  U_\gamma(\infty,s)\,
  F_{\mu\nu}(\gamma(s))\dot\gamma^\mu(s)\eta^\nu(s)\,
  U_\gamma(s,0),
\]
in the same sign convention as the preceding differential equation.  Thus a
compact nonabelian deformation inserts covariantly transported curvature along
the finite strip between the two dressings; it is not a change of the endpoint
representation at infinity.  None of these identities proves independence
under changes of the asymptotic ray, because such changes are not compact:
they modify the electric flux at infinity.  In massless QED the asymptotic
ray or Coulomb profile is tied to the charged velocity, and changing it
changes the infrared representation rather than merely the interpolating
coordinate.

\subsection{Asymptotic Wilson Lines, Flux, and Eikonal Data}

The preceding definitions identify a charged dressing as part of the
asymptotic datum.  We now spell out the elementary calculation behind this
statement.  Work in four spacetime dimensions with an abelian gauge field, and
let a massive charged particle have charge \(q\), momentum \(p^\mu\), mass
\(m>0\), and four-velocity
\[
  u^\mu=\frac{p^\mu}{m},
  \qquad u_\mu u^\mu=-1,
  \qquad p^0>0 .
\]
The future-directed straight dressing from \(x\) is represented, before the
infrared limit is taken, by the regulated current
\begin{equation}
  J_{q,u,\epsilon}^\mu(y;x)
  =
  q\int_0^\infty \dd\tau\,
  \ee^{-\epsilon\tau}u^\mu
  \delta^{(4)}(y-x-u\tau),
  \qquad \epsilon>0 .
  \label{eq:regulated-asymptotic-worldline-current}
\end{equation}
The corresponding Wilson factor is the exponential of
\(\ii\int A_\mu J^\mu\).  The factor \(\epsilon\) is not a physical photon
mass; it is an adiabatic endpoint regulator used to define the half-line
Fourier transform.  The endpoint at \(\tau=0\) is the charged field insertion,
and the endpoint at \(\tau=\infty\) is the boundary charge recorded by
\eqref{eq:charged-wilson-line-boundary-transformation}.

\paragraph{Worldline dressing and the soft eikonal denominator.}
With the Fourier convention
\[
  \widehat J^\mu(k)=\int\dd^4y\,\ee^{\ii k\cdot y}J^\mu(y),
\]
the regulated current \eqref{eq:regulated-asymptotic-worldline-current} has
\begin{equation}
  \widehat J_{q,u,\epsilon}^\mu(k;x)
  =
  q\,u^\mu\,\ee^{\ii k\cdot x}
  \frac{1}{\epsilon-\ii k\cdot u}.
  \label{eq:worldline-current-fourier-transform}
\end{equation}
For an on-shell soft photon of momentum \(k\) and physical polarization
\(e_h(k)\), the homogeneous soft factor carried by the dressing is therefore
\begin{equation}
  f_{p,\epsilon}^h(k)
  =
  q\,\frac{p\cdot e_h^*(k)}{\epsilon m-\ii p\cdot k}\,
  \ee^{\ii k\cdot x}.
  \label{eq:charged-dressed-fk-profile}
\end{equation}
After the endpoint prescription is fixed, this is the same eikonal
denominator that appears in the leading Faddeev--Kulish soft cloud and in
the leading soft-photon theorem.
Substituting \eqref{eq:regulated-asymptotic-worldline-current} gives
\[
\begin{aligned}
  \widehat J_{q,u,\epsilon}^\mu(k;x)
  &=
  q u^\mu
  \int_0^\infty\dd\tau\,
  \ee^{-\epsilon\tau}
  \ee^{\ii k\cdot(x+u\tau)}
  \\
  &=
  q u^\mu\ee^{\ii k\cdot x}
  \int_0^\infty\dd\tau\,
  \ee^{-(\epsilon-\ii k\cdot u)\tau}
  =
  q u^\mu\ee^{\ii k\cdot x}
  \frac{1}{\epsilon-\ii k\cdot u}.
\end{aligned}
\]
Since \(p^\mu=mu^\mu\), contraction with \(e_h^*(k)\) gives
\eqref{eq:charged-dressed-fk-profile}.  The sign of the infinitesimal
imaginary part depends on whether the half-line is future or past directed
and on the incoming/outgoing convention.  The load-bearing structure is the
homogeneous denominator \(p\cdot k\), not the regulator sign.

The same charged sector can be described at spacelike infinity by its
electric flux.  If \(\mathbf v=\mathbf p/p^0\) and
\(\mathbf n\in S^2\), the boosted Coulomb radial flux density is
\begin{equation}
  \mathcal E_{q,\mathbf v}(\mathbf n)
  =
  \frac{q}{4\pi}
  \frac{1-|\mathbf v|^2}{(1-\mathbf v\cdot\mathbf n)^2}.
  \label{eq:charged-boosted-coulomb-flux-density-hr}
\end{equation}
It defines the value of the limiting flux functional
\[
  \Phi_\infty(f)=
  \int_{S^2}\dd\Omega(\mathbf n)\,
  f(\mathbf n)\mathcal E_{q,\mathbf v}(\mathbf n)
\]
on a one-particle charged asymptotic datum with velocity \(\mathbf v\).

\paragraph{Velocity-labelled flux sectors.}
For \(|\mathbf v|<1\),
\begin{equation}
  \int_{S^2}\dd\Omega(\mathbf n)\,
  \mathcal E_{q,\mathbf v}(\mathbf n)
  =
  q .
  \label{eq:boosted-coulomb-flux-integral}
\end{equation}
If \(q\ne0\), the function
\(\mathcal E_{q,\mathbf v}(\mathbf n)\) determines \(\mathbf v\).  Thus
different charged velocities carry different sharp angular flux data.
Rotate coordinates so that \(\mathbf v=\beta\mathbf e_3\).  Then
\[
\begin{aligned}
  \int_{S^2}\dd\Omega\,
  \frac{1-\beta^2}{(1-\beta\cos\theta)^2}
  &=
  2\pi(1-\beta^2)
  \int_{-1}^{1}\frac{\dd z}{(1-\beta z)^2}
  \\
  &=
  2\pi(1-\beta^2)
  \left[\frac{1}{\beta(1-\beta z)}\right]_{z=-1}^{z=1}
  =
  4\pi ,
\end{aligned}
\]
which proves \eqref{eq:boosted-coulomb-flux-integral}.  For \(q\ne0\), the
maximum of \(\mathcal E_{q,\mathbf v}\) occurs in the direction of
\(\mathbf v\) when \(\mathbf v\ne0\), while the ratio of the maximum to the
minimum is
\[
  \left(\frac{1+|\mathbf v|}{1-|\mathbf v|}\right)^2 .
\]
This ratio determines \(|\mathbf v|\), and the direction of the maximum
determines the direction of \(\mathbf v\).  The case \(\mathbf v=0\) is the
unique rotationally invariant density.

\paragraph{Neutral boundary charge versus angular flux sector.}
The boundary-charge condition
\eqref{eq:dressed-correlator-boundary-charge-selection} is the zero-mode of
a sharper asymptotic datum.  For an oriented list of charged asymptotic
coordinates
\[
  I=((\epsilon_1,q_1,\mathbf v_1),\ldots,
     (\epsilon_N,q_N,\mathbf v_N)),
  \qquad \epsilon_a=\pm1,
\]
define the signed angular flux profile
\begin{equation}
  \mathcal E_I(\mathbf n)
  =
  \sum_{a=1}^{N}\epsilon_a\,
  \mathcal E_{q_a,\mathbf v_a}(\mathbf n).
  \label{eq:charged-multiparticle-angular-flux-profile}
\end{equation}
By \eqref{eq:boosted-coulomb-flux-integral},
\[
  \int_{S^2}\dd\Omega(\mathbf n)\,\mathcal E_I(\mathbf n)
  =
  \sum_{a=1}^{N}\epsilon_a q_a .
\]
Thus a vacuum dressed correlator can obey the ordinary boundary-charge Ward
identity while still having nonzero angle-dependent flux data.  For instance,
two opposite charges with the same velocity have
\(\mathcal E_I(\mathbf n)=0\) pointwise.  Two opposite charges with distinct
velocities have zero total boundary charge but, unless the velocity labels
coincide, the profile \(\mathcal E_I\) is not the zero function.  The
charged LSZ datum therefore contains more information than the integer or
real total charge: it contains the limiting angular Gauss-law profile, or an
equivalent large-gauge-charge functional.  A same-flux dressing change is a
finite coordinate change of the type described above; a change of
\(\mathcal E_I\) changes the asymptotic charged sector and belongs to the
open Haag--Ruelle problem, not to a harmless path-independence argument.

\paragraph{Flux-sector projection in dressed LSZ.}
The same distinction has to be imposed before one takes simultaneous
external-shell residues.  Let \(\mathfrak F\) and \(\mathfrak F'\) be
characters of the abelian algebra generated by the large-sphere flux
operators \(\Phi_\infty(f)\), so that \(\mathfrak F(f)\) is the angular
Gauss-law flux of the initial charged sector tested against \(f\).  For an
oriented external list \(I\) as in
\eqref{eq:charged-multiparticle-angular-flux-profile}, the dressed LSZ
residue is block diagonal in these flux characters:
\begin{equation}
  P_{\mathfrak F'}\operatorname{Res}_{I}^{\rm LSZ}(G)P_{\mathfrak F}
  =
  0
  \quad\text{unless}\quad
  \mathfrak F'(f)-\mathfrak F(f)
  =
  \int_{S^2}\dd\Omega(\mathbf n)\,
  f(\mathbf n)\mathcal E_I(\mathbf n)
  \quad\text{for all }f .
  \label{eq:dressed-lsz-flux-sector-selection}
\end{equation}
Taking \(f=1\) gives the boundary-charge Ward rule
\eqref{eq:dressed-correlator-boundary-charge-selection}; the nonconstant
test functions are the extra large-gauge charges carried by the angular flux.
Thus a vacuum-to-vacuum dressed correlator requires
\(\mathcal E_I=0\) as a function on the sphere; vanishing of
\(\int_{S^2}\mathcal E_I\) only checks the charge zero mode.  A
charge-neutral pair with distinct charged velocities is therefore either a
matrix element between different flux sectors, or it vanishes after
projection to the vacuum flux block.

In the finite-regulator isolated-shell notation this means that the external
overlap maps and their left inverses must be chosen inside one flux block:
\begin{equation}
  z(p)=\bigoplus_{\mathfrak F} z_{\mathfrak F}(p),
  \qquad
  L(p)=\bigoplus_{\mathfrak F} L_{\mathfrak F}(p),
  \qquad
  L_{\mathfrak F}(p)z_{\mathfrak F}(p)=\mathbf 1_{E_{p,\mathfrak F}} .
  \label{eq:dressed-lsz-flux-block-overlap}
\end{equation}
Finite same-flux coordinate changes act inside a single summand of
\eqref{eq:dressed-lsz-flux-block-overlap} and cancel against the transformed
left inverse as in \eqref{eq:dressed-lsz-left-inverse-residue}.  A matrix
that mixes two different angular-flux characters is not a dressing-coordinate
change on one charged shell; it changes the sector on which the large-time
wave map is supposed to land.  If a screening regulator collapses the flux
characters, this block statement reduces to the ordinary finite-coordinate
algebra.  Without such a collapse it is a superselection projection that has
to precede the scalar LSZ normalization.

\subsection{Soft Coherent Norm and Velocity Superselection}

The preceding flux density is visible already in the finite-cutoff soft
photon Fock calculation.  The calculation is not a construction of massless
QED scattering; it is a precise diagnostic of why a single ordinary photon
Fock representation cannot contain sharp charged particles of all velocities
as normal vectors after the infrared cutoff is removed.

Fix \(0<\lambda<\Lambda<\infty\).  The finite soft-photon one-particle space
is
\[
  \mathfrak h_{\lambda,\Lambda}
  =
  L^2\!\left(
    [\lambda,\Lambda]\times S^2,
    \frac{r\,\dd r\,\dd\Omega(\mathbf n)}{2(2\pi)^3};
    \mathbb C^2
  \right),
\]
where \(k=(r,r\mathbf n)\) is null and the two complex components label a
measurable orthonormal transverse polarization frame
\(\epsilon_h(\mathbf n)\), \(h=1,2\), satisfying
\(\mathbf n\cdot\epsilon_h(\mathbf n)=0\).  The polarization frame is a
coordinate choice; the sum over \(h\) below is the transverse projector
\[
  P_{\mathbf n}^{ij}=\delta^{ij}-n^in^j .
\]
For a massive charge \(q\) with velocity \(\mathbf v\), \(|\mathbf v|<1\), the
homogeneous soft part of the worldline dressing
\eqref{eq:charged-dressed-fk-profile} is represented by
\begin{equation}
  F_{q,\mathbf v,\lambda,\Lambda}^{h}(r,\mathbf n)
  =
  -q\,\frac{\mathbf v\cdot\epsilon_h(\mathbf n)}
           {r(1-\mathbf v\cdot\mathbf n)}
  \mathbf 1_{\lambda<r<\Lambda}.
  \label{eq:charged-soft-coherent-profile-velocity}
\end{equation}
The mass and Lorentz factor cancel because
\[
  \frac{p\cdot e_h(k)}{p\cdot k}
  =
  -\frac{\mathbf v\cdot\epsilon_h(\mathbf n)}
          {r(1-\mathbf v\cdot\mathbf n)}
\]
for \(e_h^0=0\), \(p^\mu=m\gamma(1,\mathbf v)\), and the mostly-plus
Minkowski product.

\begin{proposition}[Velocity separation in the soft coherent norm]
\label{prop:soft-coherent-velocity-separation}
For two velocities \(\mathbf v,\mathbf w\in B_1(0)\subset\mathbb R^3\) and the
same nonzero charge \(q\),
\begin{equation}
  \left\|
    F_{q,\mathbf v,\lambda,\Lambda}
    -
    F_{q,\mathbf w,\lambda,\Lambda}
  \right\|_{\mathfrak h_{\lambda,\Lambda}}^2
  =
  \frac{q^2}{2(2\pi)^3}
  \log\frac{\Lambda}{\lambda}\,
  \mathcal A(\mathbf v,\mathbf w),
  \label{eq:soft-coherent-velocity-log-norm}
\end{equation}
where
\begin{equation}
  \mathcal A(\mathbf v,\mathbf w)
  =
  \int_{S^2}\dd\Omega(\mathbf n)\,
  \left|
    P_{\mathbf n}
    \left(
      \frac{\mathbf v}{1-\mathbf v\cdot\mathbf n}
      -
      \frac{\mathbf w}{1-\mathbf w\cdot\mathbf n}
    \right)
  \right|^2 .
  \label{eq:soft-coherent-angular-coefficient}
\end{equation}
The coefficient obeys \(\mathcal A(\mathbf v,\mathbf w)\ge0\), and
\(\mathcal A(\mathbf v,\mathbf w)=0\) if and only if
\(\mathbf v=\mathbf w\).
\end{proposition}

\begin{proof}
The radial dependence of
\eqref{eq:charged-soft-coherent-profile-velocity} is \(r^{-1}\).  Therefore
the measure \(r\,\dd r\,\dd\Omega/(2(2\pi)^3)\) gives the logarithm
\[
  \int_\lambda^\Lambda \frac{\dd r}{r}=\log\frac{\Lambda}{\lambda}.
\]
After summing over the two physical polarizations,
\[
  \sum_{h=1}^2
  \left|
    \left(
      \frac{\mathbf v}{1-\mathbf v\cdot\mathbf n}
      -
      \frac{\mathbf w}{1-\mathbf w\cdot\mathbf n}
    \right)
    \cdot\epsilon_h(\mathbf n)
  \right|^2
  =
  \left|
    P_{\mathbf n}
    \left(
      \frac{\mathbf v}{1-\mathbf v\cdot\mathbf n}
      -
      \frac{\mathbf w}{1-\mathbf w\cdot\mathbf n}
    \right)
  \right|^2 ,
\]
which proves \eqref{eq:soft-coherent-velocity-log-norm}.  The integrand in
\(\mathcal A(\mathbf v,\mathbf w)\) is the squared norm of a transverse
vector for each \(\mathbf n\), and the sphere measure is positive, so
\(\mathcal A(\mathbf v,\mathbf w)\ge0\).

It remains to characterize the zero set.  If
\(\mathcal A(\mathbf v,\mathbf w)=0\), continuity implies that the displayed
transverse projection vanishes for every \(\mathbf n\).  Suppose first that
\(\mathbf v\ne0\).  Put
\(\mathbf n=\mathbf v/|\mathbf v|\).  The projection of \(\mathbf v\) onto
the plane orthogonal to \(\mathbf n\) is zero, hence the projection of
\(\mathbf w\) onto that plane must also vanish; \(\mathbf w\) is parallel to
\(\mathbf v\).  Writing
\(\mathbf v=\beta\mathbf e_3\) and
\(\mathbf w=\alpha\mathbf e_3\), and choosing any
\(\mathbf n\) not parallel to \(\mathbf e_3\), the vanishing condition becomes
\[
  \frac{\beta}{1-\beta n_3}
  =
  \frac{\alpha}{1-\alpha n_3}.
\]
Thus \(\beta=\alpha\).  If \(\mathbf v=0\), the same condition says that the
transverse projection of \(\mathbf w\) vanishes for every \(\mathbf n\); this
forces \(\mathbf w=0\).  Hence the coefficient vanishes exactly on the
diagonal \(\mathbf v=\mathbf w\).
\end{proof}

Let \(\mathcal E(F)\) denote the normalized bosonic coherent vector in the
finite-cutoff photon Fock space,
\[
  \mathcal E(F)
  =
  \ee^{-\|F\|^2/2}
  \bigoplus_{n\ge0}\frac{F^{\otimes_s n}}{\sqrt{n!}} .
\]
Summing the Fock inner products gives
\[
  \left|
    \langle \mathcal E(F),\mathcal E(G)\rangle
  \right|
  =
  \exp\!\left\{-\frac12\|F-G\|^2\right\}.
\]
Therefore Proposition~\ref{prop:soft-coherent-velocity-separation} implies
that, for \(\mathbf v\ne\mathbf w\),
\[
  \left|
    \left\langle
      \mathcal E(F_{q,\mathbf v,\lambda,\Lambda}),
      \mathcal E(F_{q,\mathbf w,\lambda,\Lambda})
    \right\rangle
  \right|
  =
  \exp\!\left\{
    -\frac{q^2}{4(2\pi)^3}
    \log\frac{\Lambda}{\lambda}\,
    \mathcal A(\mathbf v,\mathbf w)
  \right\}
  \longrightarrow0
  \qquad(\lambda\downarrow0).
\]
This is the finite-Fock-space calculation behind velocity-labelled charged
sectors.  It is stronger than saying that a bare electron emits many soft
photons: different charged velocities determine different infrared coherent
directions.  A Haag--Ruelle replacement for massless charged sectors must
therefore build its asymptotic representation with the charged velocity and
flux label included from the start.

\begin{controlledapproximation}[Ray--velocity matching for charged external sectors]
\label{ca:charged-ray-velocity-matching}
In the finite-cutoff abelian model, an external massive charged packet with
velocity support \(V\subset B_1(0)\) must be dressed by an asymptotic ray or
Coulomb profile assigned over the same velocity cell.  Write this assignment
as \(\Gamma(\mathbf v)\), and let its homogeneous soft profile be
\(F_{q,\Gamma(\mathbf v),\lambda,\Lambda}\).  The admissible matching
condition is
\[
  F_{q,\Gamma(\mathbf v),\lambda,\Lambda}
  =
  F_{q,\mathbf v,\lambda,\Lambda}
  +
  \Delta_{\mathbf v,\lambda,\Lambda},
  \qquad
  \sup_{\mathbf v\in V}
  \limsup_{\lambda\downarrow0}
  \|\Delta_{\mathbf v,\lambda,\Lambda}\|_{\mathfrak h_{\lambda,\Lambda}}
  <\infty ,
\]
for each fixed ultraviolet window \(\Lambda\).  Thus compact changes of the
path, or finite Hilbert soft reparametrizations inside one flux block, are
allowed; replacing the velocity profile by a different asymptotic ray is not.

Indeed, if the same packet velocity \(\mathbf v\) is divided by a dressing
profile built from a different velocity \(\mathbf w\), then
Proposition~\ref{prop:soft-coherent-velocity-separation} gives
\[
  \|F_{q,\mathbf v,\lambda,\Lambda}
     -F_{q,\mathbf w,\lambda,\Lambda}\|^2
  =
  \frac{q^2}{2(2\pi)^3}
  \log\frac{\Lambda}{\lambda}\,
  \mathcal A(\mathbf v,\mathbf w),
  \qquad
  \mathcal A(\mathbf v,\mathbf w)>0 .
\]
The mismatch is therefore a new infrared sector, not a finite coordinate
change of the external charged shell.  In the soft-subtracted boundary-value
notation this says that the factor removed from a leg with momentum \(p\)
must be the one with \(p^\mu=m u^\mu\):
\[
  \frac{p\cdot e_h(k)}{p\cdot k-\ii0}
  =
  \frac{u\cdot e_h(k)}{u\cdot k-\ii0}.
\]
Using another ray \(u'\) leaves a residual homogeneous \(1/r\) soft profile.
In the Dollard comparison the same mistake appears as a nonzero limiting
coefficient error \(\delta\kappa\), hence as an unremoved
\(\delta\kappa\,\dd t/t\) phase derivative.  The large-time scalar products
then retain the factor \(\exp(\ii\delta\kappa\log t)\), or the finite-cutoff
coherent overlap vanishes as the infrared cutoff is removed.

Consequently a dressed LSZ coordinate is not specified by charge and an
arbitrary Wilson-line path alone.  For each charged external packet one must
record the velocity cell, the associated angular flux character, and the
matched asymptotic ray or Coulomb tail before applying the left inverse in
\eqref{eq:dressed-lsz-flux-block-overlap}.  Negative tests are now precise:
a fixed ray used for all charged momenta, a charge-only soft factor, or an
\(R\)-matrix that mixes different asymptotic rays cannot be repaired by the
same-flux coordinate algebra of \eqref{eq:dressed-lsz-left-inverse-residue}.
\end{controlledapproximation}

\subsection{Weyl Implementation and the Failure of a Common Photon Fock Space}

The preceding norm calculation has an operator-algebraic meaning.  For
\(0<\Lambda<\infty\), let \(\mathfrak h_{0,\Lambda}\) be the photon
one-particle space obtained by replacing the interval
\([\lambda,\Lambda]\) in \(\mathfrak h_{\lambda,\Lambda}\) by
\((0,\Lambda]\).  For each \(\lambda>0\),
\(\mathfrak h_{\lambda,\Lambda}\) is a closed subspace of
\(\mathfrak h_{0,\Lambda}\) by extension by zero on \((0,\lambda]\).  Let
\[
  \mathcal F_{0,\Lambda}=\mathcal F_s(\mathfrak h_{0,\Lambda})
\]
be the corresponding fixed photon Fock space, and let
\[
  W(f)=\exp\!\left(a^\dagger(f)-a(f)\right),
  \qquad f\in\mathfrak h_{0,\Lambda},
\]
be the Weyl operator.  With the real symplectic form
\[
  \sigma(f,g)=2\,\operatorname{Im}\langle f,g\rangle_{\mathfrak h},
\]
the Weyl relations are
\begin{equation}
  W(f)W(g)=
  \exp\!\left(-\frac{\ii}{2}\sigma(f,g)\right)W(f+g),
  \qquad
  W(f)^*=W(-f),
  \label{eq:charged-sector-weyl-relations}
\end{equation}
and the Fock vacuum state is
\[
  \omega_0(W(f))=\langle\Omega,W(f)\Omega\rangle
  =
  \exp\!\left(-\frac12\|f\|_{\mathfrak h}^2\right).
\]
The coherent state with soft profile \(F\) is
\[
  \omega_F(A)=
  \langle \mathcal E(F),A\mathcal E(F)\rangle
  =
  \langle W(F)\Omega,A W(F)\Omega\rangle .
\]
On Weyl generators this gives
\begin{equation}
  \omega_F(W(f))
  =
  \exp\!\left(\ii\sigma(F,f)\right)
  \exp\!\left(-\frac12\|f\|_{\mathfrak h}^2\right).
  \label{eq:coherent-weyl-state-characteristic-functional}
\end{equation}
Thus the soft cloud changes the representation by adding a linear phase on
the Weyl algebra.  At finite cutoff this phase is inner, implemented by the
unitary \(W(F)\).  The infrared question is whether the implementing unitaries
have a limit on one fixed Fock representation as \(\lambda\downarrow0\).

\begin{proposition}[No common Fock implementer for distinct charged velocities]
\label{prop:no-common-fock-implementer-distinct-velocities}
Fix \(q\ne0\), \(0<\Lambda<\infty\), and two distinct velocities
\(\mathbf v,\mathbf w\in B_1(0)\).  Put
\[
  D_{\lambda,\Lambda}
  =
  F_{q,\mathbf v,\lambda,\Lambda}
  -
  F_{q,\mathbf w,\lambda,\Lambda}.
\]
Then \(\|D_{\lambda,\Lambda}\|\to\infty\) as
\(\lambda\downarrow0\).  Consequently the finite-cutoff Weyl implementers
\(W(D_{\lambda,\Lambda})\) have no strong operator limit, and no nonzero weak
operator limit, on the fixed photon Fock representation.
\end{proposition}

\begin{proof}
The divergence of \(\|D_{\lambda,\Lambda}\|\) is exactly
Proposition~\ref{prop:soft-coherent-velocity-separation}: for
\(\mathbf v\ne\mathbf w\),
\[
  \|D_{\lambda,\Lambda}\|^2
  =
  \frac{q^2}{2(2\pi)^3}
  \log\frac{\Lambda}{\lambda}\,
  \mathcal A(\mathbf v,\mathbf w),
  \qquad
  \mathcal A(\mathbf v,\mathbf w)>0 .
\]
It remains to translate this divergence into an operator statement.  Let
\(\mathcal D_{\rm fin}\) be the dense subspace of finite-particle vectors.
Matrix elements of \(W(D)\) between vectors in
\(\mathcal D_{\rm fin}\) are finite sums of polynomials in the finitely many
inner products \(\langle f_i,D\rangle\) and their complex conjugates,
multiplied by the vacuum Gaussian factor
\[
  \langle\Omega,W(D)\Omega\rangle
  =
  \exp\!\left(-\frac12\|D\|^2\right).
\]
This follows by commuting \(a(f)\) and \(a^\dagger(f)\) through \(W(D)\), or
equivalently by differentiating the Weyl relation
\eqref{eq:charged-sector-weyl-relations} with respect to finite test
directions.  Polynomial growth cannot compensate the Gaussian decay.  Hence
for every \(\xi,\eta\in\mathcal D_{\rm fin}\),
\[
  \langle\xi,W(D_{\lambda,\Lambda})\eta\rangle
  \longrightarrow0
  \qquad(\lambda\downarrow0).
\]
The operators \(W(D_{\lambda,\Lambda})\) are unitary and therefore uniformly
bounded, so density of \(\mathcal D_{\rm fin}\) gives weak convergence to zero
on all Fock vectors.  A strong limit of unitaries, if it existed, would have
the same weak limit and would therefore be the zero operator; but a strong
limit of unitaries is an isometry on every vector whose norm limit exists:
\[
  \|U\eta\|=\lim_{\lambda\downarrow0}
  \|W(D_{\lambda,\Lambda})\eta\|=\|\eta\|.
\]
The zero operator is not an isometry.  Thus no strong limit exists.
\end{proof}

\begin{definition}[Hilbert-equivalent soft dressings]
\label{def:hilbert-equivalent-soft-dressings}
Fix the ultraviolet cutoff \(\Lambda\) and use the photon one-particle space
\(\mathfrak h_{0,\Lambda}\) introduced above.  Let
\[
  F_{\lambda},G_{\lambda}\in\mathfrak h_{\lambda,\Lambda}
  \subset\mathfrak h_{0,\Lambda},
  \qquad 0<\lambda<\Lambda,
\]
be two infrared-regulated soft profiles for the same charged insertion.  They
are called Hilbert-equivalent at the cutoff \(\Lambda\) if there exists a
vector \(\Delta\in\mathfrak h_{0,\Lambda}\) such that
\begin{equation}
  \lim_{\lambda\downarrow0}
  \|G_{\lambda}-F_{\lambda}-\Delta\|_{\mathfrak h_{0,\Lambda}}
  =
  0 .
  \label{eq:soft-profile-hilbert-equivalence}
\end{equation}
The vector \(\Delta\) is the finite soft-coordinate change between the two
dressings.  This definition is deliberately a Hilbert-space definition: it
does not identify dressings whose difference changes the limiting Gauss-law
flux, because such a difference is not represented by a vector of
\(\mathfrak h_{0,\Lambda}\).
\end{definition}

\begin{proposition}[Hilbert soft changes are inner Weyl coordinate changes]
\label{prop:hilbert-soft-changes-inner-weyl-coordinate}
Let \(F_{\lambda}\) and \(G_{\lambda}\) be Hilbert-equivalent soft profiles in
the sense of Definition~\ref{def:hilbert-equivalent-soft-dressings}, and put
\[
  \Delta_{\lambda}=G_{\lambda}-F_{\lambda}.
\]
For every finite Weyl polynomial \(A\) in the generators \(W(f)\),
\(f\in\mathfrak h_{0,\Lambda}\),
\begin{equation}
  \omega_{G_{\lambda}}(A)
  -
  \omega_{F_{\lambda}}\!\left(
    W(\Delta_{\lambda})^* A W(\Delta_{\lambda})
  \right)
  =
  0 ,
  \label{eq:finite-cutoff-soft-coordinate-inner-equality}
\end{equation}
and the implementing unitaries have the strong limit
\[
  W(\Delta_{\lambda})
  \longrightarrow
  W(\Delta)
  \qquad\text{on }\mathcal F_{0,\Lambda}.
\]
Consequently
\[
  |\langle\mathcal E(F_{\lambda}),\mathcal E(G_{\lambda})\rangle|
  =
  \exp\!\left(-\frac12\|\Delta_{\lambda}\|^2\right)
  \longrightarrow
  \exp\!\left(-\frac12\|\Delta\|^2\right)>0 .
\]
Thus a Hilbert-equivalent soft change is a finite reparametrization of the
same infrared representation, not a change of charged sector.  The velocity
change in
Proposition~\ref{prop:no-common-fock-implementer-distinct-velocities} fails
this criterion because
\(\|F_{q,\mathbf v,\lambda,\Lambda}
-F_{q,\mathbf w,\lambda,\Lambda}\|\to\infty\) for
\(\mathbf v\ne\mathbf w\).
\end{proposition}

\begin{proof}
The equality \eqref{eq:finite-cutoff-soft-coordinate-inner-equality} is first
checked on one Weyl generator.  The Weyl relations give
\[
  W(\Delta_{\lambda})^*W(f)W(\Delta_{\lambda})
  =
  \exp\!\left(\ii\sigma(\Delta_{\lambda},f)\right)W(f),
\]
and hence, using
\eqref{eq:coherent-weyl-state-characteristic-functional},
\[
\begin{aligned}
  \omega_{F_{\lambda}}\!\left(
    W(\Delta_{\lambda})^*W(f)W(\Delta_{\lambda})
  \right)
  &=
  \exp\!\left(\ii\sigma(\Delta_{\lambda},f)\right)
  \exp\!\left(\ii\sigma(F_{\lambda},f)\right)
  \exp\!\left(-\frac12\|f\|^2\right)
\\
  &=
  \exp\!\left(\ii\sigma(G_{\lambda},f)\right)
  \exp\!\left(-\frac12\|f\|^2\right)
  =
  \omega_{G_{\lambda}}(W(f)).
\end{aligned}
\]
Linearity and multiplication by the Weyl relations extend this identity to
finite Weyl polynomials.

It remains to prove the strong convergence of the implementers.  The coherent
vectors
\[
  \mathcal E(K)
  =
  \ee^{-\|K\|^2/2}
  \bigoplus_{n\ge0}\frac{K^{\otimes_s n}}{\sqrt{n!}},
  \qquad K\in\mathfrak h_{0,\Lambda},
\]
span a dense subspace of \(\mathcal F_{0,\Lambda}\).  On such vectors,
\[
  W(H)\mathcal E(K)
  =
  \exp\!\left(-\frac{\ii}{2}\sigma(H,K)\right)
  \mathcal E(K+H),
\]
because \(\mathcal E(K)=W(K)\Omega\) and
\eqref{eq:charged-sector-weyl-relations} gives
\(W(H)W(K)=\exp[-\ii\sigma(H,K)/2]W(H+K)\).  If \(H_n\to H\) in
\(\mathfrak h_{0,\Lambda}\), the displayed formula and the coherent-vector
inner product
\[
  \langle\mathcal E(K_1),\mathcal E(K_2)\rangle
  =
  \exp\!\left(
    -\frac12\|K_1\|^2-\frac12\|K_2\|^2+\langle K_1,K_2\rangle
  \right)
\]
show that \(W(H_n)\mathcal E(K)\to W(H)\mathcal E(K)\) in norm.  Since all
\(W(H_n)\) and \(W(H)\) are unitary, density upgrades convergence on coherent
vectors to strong convergence on the whole Fock space.  Applying this to
\(H_n=\Delta_{\lambda_n}\) along every sequence
\(\lambda_n\downarrow0\) proves
\[
  W(\Delta_{\lambda})\to W(\Delta)
  \quad\text{strongly}.
\]
The overlap formula follows by summing the Fock inner products of coherent
vectors, and the limiting overlap is nonzero because
\(\Delta\in\mathfrak h_{0,\Lambda}\) has finite norm.  The last assertion is
then exactly
Proposition~\ref{prop:no-common-fock-implementer-distinct-velocities}.
\end{proof}

Proposition~\ref{prop:no-common-fock-implementer-distinct-velocities} is a
precise version of the representation-theoretic obstruction.  At every
positive infrared cutoff, changing the soft cloud is an ordinary Weyl
displacement on Fock space.  Removing the cutoff separates two cases.  A
Hilbert-equivalent change remains an inner coordinate change by
Proposition~\ref{prop:hilbert-soft-changes-inner-weyl-coordinate}.  A change
of charged velocity escapes to infinity in the one-particle norm and is not
implemented on the original Fock representation.  Therefore a charged
asymptotic Hilbert space must be organized as a direct integral or measurable
field of infrared representations labelled by the limiting flux data, not as
a single photon Fock space tensored with a Wigner one-particle electron
space.  This is still not the missing Haag--Ruelle theorem, but it is one of
the structural inputs that theorem must contain.

The flux statement and the eikonal statement are two aspects of the same
nonlocal charged datum.  The Wilson line fixes how the charged insertion
transforms under boundary gauge transformations; the boosted Coulomb field
records the corresponding Gauss-law flux at spacelike infinity; the
half-line Fourier transform gives the soft denominator used by perturbative
dressed-state constructions.  A local observable cannot change
\(\Phi_\infty(f)\), so a local observable cannot replace this noncompact
dressing in a Haag--Ruelle construction.  Conversely, specifying only the
formal denominator \(p\cdot k\) is not yet a scattering theorem: a full
charged-sector Haag--Ruelle theory must still construct large-time limits in
the representation labelled by the flux density
\eqref{eq:charged-boosted-coulomb-flux-density-hr}.

\subsection{Velocity-Fibered Soft Representations}

The preceding propositions rule out one tempting but incorrect construction:
one cannot keep a single photon Fock representation and implement all charged
velocity changes by limiting Weyl unitaries.  The minimal replacement is a
fibered infrared representation.  A scattering theory must still construct
large-time limits, while this representation gives the correct
representation-theoretic target for such a construction in a
nonconfining charged sector.

This qualification is essential.  The charged-sector construction discussed
here is not a proposed scattering theory for colored quarks in a confining
four-dimensional Yang--Mills theory.  It applies only when the theory has
physical charged superselection sectors with charged asymptotic data: for
example electric charge in a nonconfining abelian gauge theory, or suitable
unconfined charges on a Coulomb branch or in a deconfined regime after the
infrared and representation-theoretic hypotheses have been stated.  In a
confining theory the gauge-invariant Wilson-line operator ending at infinity
creates a state with a macroscopic flux string or fails to have a
finite-energy charged-particle limit; the physical massive one-particle
states, if present, are instead neutral hadrons or glueballs created by local
gauge-invariant operators.  Ordinary Haag--Ruelle theory is then the correct
starting point for those neutral particles.

Fix \(q\ne0\) and \(0<\Lambda<\infty\).  Let
\(\mathfrak h_{\rm reg}\subset\mathfrak h_{0,\Lambda}\) be a complex linear
subspace of infrared-regular photon test functions such that, for every
\(\mathbf v\in B_1(0)\), the limit
\begin{equation}
  \ell_{\mathbf v}(f)
  =
  \lim_{\lambda\downarrow0}
  \sigma(F_{q,\mathbf v,\lambda,\Lambda},f)
  \label{eq:velocity-soft-character-limit}
\end{equation}
exists for all \(f\in\mathfrak h_{\rm reg}\).  For example, the condition is
satisfied on smooth compactly supported one-photon wave packets whose radial
support is bounded away from \(r=0\).  Larger infrared test spaces require an
explicit condition on the radial behavior of \(f\); the definition above
records exactly the condition needed for the soft phase to be a finite
functional.

Let \(\mathfrak W_{\rm reg}\) be the Weyl \(*\)-algebra generated by
\(W(f)\), \(f\in\mathfrak h_{\rm reg}\), with relations
\eqref{eq:charged-sector-weyl-relations}.  The charged velocity
\(\mathbf v\) defines the state
\begin{equation}
  \omega_{\mathbf v}(W(f))
  =
  \exp\!\left(\ii\ell_{\mathbf v}(f)\right)
  \exp\!\left(-\frac12\|f\|^2_{\mathfrak h_{0,\Lambda}}\right),
  \qquad f\in\mathfrak h_{\rm reg}.
  \label{eq:velocity-fiber-soft-state}
\end{equation}
Let \((\mathcal H_{\mathbf v},\pi_{\mathbf v},\Omega_{\mathbf v})\) be the
GNS representation of \(\omega_{\mathbf v}\).  If \(\nu\) is a Borel measure
on \(B_1(0)\) for which the field
\(\mathbf v\mapsto(\mathcal H_{\mathbf v},\pi_{\mathbf v})\) has been given a
measurable structure, define
\begin{equation}
  \mathcal H_{q,\nu}^{\rm soft}
  =
  \int_{B_1(0)}^{\oplus}\mathcal H_{\mathbf v}\,\dd\nu(\mathbf v),
  \qquad
  \bigl(\pi_{q,\nu}(A)\Psi\bigr)(\mathbf v)
  =
  \pi_{\mathbf v}(A)\Psi(\mathbf v).
  \label{eq:charged-soft-direct-integral-representation}
\end{equation}

\begin{remark}[Soft photon algebra preserves the velocity fibers]
\label{rem:soft-weyl-algebra-preserves-velocity-fibers}
In the direct-integral representation
\eqref{eq:charged-soft-direct-integral-representation}, let
\(P_E\) be multiplication by the characteristic function of a Borel set
\(E\subset B_1(0)\).  Then
\begin{equation}
  [P_E,\pi_{q,\nu}(A)]=0
  \qquad
  \text{for all }A\in\mathfrak W_{\rm reg}.
  \label{eq:velocity-projection-commutes-soft-algebra}
\end{equation}
Consequently, if \(E\cap E'=\varnothing\), then
\begin{equation}
  \langle P_E\xi,\pi_{q,\nu}(A)P_{E'}\eta\rangle=0
  \qquad
  \text{for all }\xi,\eta\in\mathcal H_{q,\nu}^{\rm soft},
  \quad A\in\mathfrak W_{\rm reg}.
  \label{eq:soft-algebra-no-velocity-transition}
\end{equation}
Thus the soft photon Weyl algebra is diagonal in the charged velocity label.
Any charged scattering operator that changes the massive particle momentum
must include additional charged-sector dynamics supplied by operators beyond
the soft photon algebra alone.
Indeed, by the definition of the direct-integral action, for
\(\Psi\in\mathcal H_{q,\nu}^{\rm soft}\),
\[
  (P_E\pi_{q,\nu}(A)\Psi)(\mathbf v)
  =
  \mathbf 1_E(\mathbf v)\pi_{\mathbf v}(A)\Psi(\mathbf v)
  =
  (\pi_{q,\nu}(A)P_E\Psi)(\mathbf v)
\]
for \(\nu\)-almost every \(\mathbf v\).  If \(E\cap E'=\varnothing\), then
\(P_EP_{E'}=0\), and therefore
\[
  \langle P_E\xi,\pi_{q,\nu}(A)P_{E'}\eta\rangle
  =
  \langle \xi,P_E\pi_{q,\nu}(A)P_{E'}\eta\rangle
  =
  \langle \xi,\pi_{q,\nu}(A)P_EP_{E'}\eta\rangle
  =
  0 .
  \]
\end{remark}

This fibered construction is the algebraic form of the velocity
superselection conclusion.  It also states a precise target for the still
missing charged Haag--Ruelle theorem: the theorem must construct incoming and
outgoing charged wave operators whose asymptotic photon part lives in a
measurable field of infrared representations labelled by the Gauss-law flux,
and whose hard charged part supplies the off-diagonal momentum-changing
dynamics.  Treating the charged soft cloud as a removable unitary dressing on
one Fock space would erase exactly the sector information proved above.

\subsection{Flux-Resolved Amplitudes and Detector-Inclusive Probabilities}

The preceding construction identifies the Hilbert-space target of a charged
LSZ theorem.  The detector observable is a further inclusive probability
built from that target.  A dressed LSZ residue in a fixed angular-flux sector
gives a hard charged coefficient; a finite-resolution experiment measures a
probability after unresolved soft photons and detector-indistinguishable flux
cells have been treated inclusively.  These are different operations, and
confusing them erases the physics of charged sectors.

Let
\[
  \mathcal M^{\mathfrak F',\mathfrak F}_{\beta\alpha}
\]
denote the flux-resolved charged coefficient extracted from the dressed LSZ
interface, with incoming and outgoing hard charged data \(\alpha,\beta\) and
initial and final angular-flux characters \(\mathfrak F,\mathfrak F'\).  The
coefficient is defined only when the flux-selection rule
\eqref{eq:dressed-lsz-flux-sector-selection} is satisfied.  For fixed hard
data, write the leading soft mismatch profile in the energy window
\(\lambda<\omega<M\) as
\[
  F_{\beta\alpha}^{h}(k)
  =
  \sum_{a\in\alpha\cup\beta}
  \eta_a q_a\,
  \frac{p_a\cdot e_h(k)}{p_a\cdot k},
  \qquad
  \eta_a=
  \begin{cases}
    +1,& a\text{ outgoing},\\
    -1,& a\text{ incoming}.
  \end{cases}
\]
The Ward identity for a physical hard transition is the statement that this
soft factor is transverse under \(e_h(k)\mapsto e_h(k)+c\,k\).  Thus it is the
same eikonal datum that appears in the inclusive calculation of
Chapter~\ref{ch:volume-ii-infrared-inclusive-qed}, but here it is attached to
a flux-resolved dressed coefficient rather than to a bare charged
Fock-space amplitude.

Let \(P_{\lambda,E_T}\) project the soft one-photon space to
\(\lambda<\omega<E_T\), and let \(P_{E_T,M}\) project it to
\(E_T<\omega<M\).  At leading homogeneous soft order,
\[
  \|P_{\lambda,E_T}F_{\beta\alpha}\|^2
  =
  A_{\beta\alpha}\log\frac{E_T}{\lambda},
  \qquad
  \|P_{E_T,M}F_{\beta\alpha}\|^2
  =
  A_{\beta\alpha}\log\frac{M}{E_T},
\]
with the same nonnegative angular coefficient \(A_{\beta\alpha}\) as in the
Bloch--Nordsieck factor
\eqref{eq:qed-bloch-nordsieck-inclusive-factorization}.  Finite soft
matching terms are part of the hard factor and detector function; the
displayed logarithms are the universal part.

\begin{proposition}[Detector-inclusive soft factor from a flux-resolved LSZ coefficient]
\label{prop:dressed-lsz-detector-inclusive-soft-factor}
At positive infrared cutoff \(\lambda\), the leading-soft probability for
the flux-resolved charged coefficient
\(\mathcal M^{\mathfrak F',\mathfrak F}_{\beta\alpha}\), with no photons
resolved above \(E_T\) inside the soft window \(E_T<\omega<M\), is
\begin{equation}
\begin{aligned}
  P_{\beta\alpha}^{(\lambda)}
  (\mathfrak F',\mathfrak F;E_T,M)
  &=
  \left|
    \mathcal M^{\mathfrak F',\mathfrak F}_{\beta\alpha}
  \right|^2
  \exp\!\left\{
    -\|F_{\beta\alpha}\|_{\lambda,M}^2
  \right\}
  \sum_{N=0}^{\infty}
  \frac{
    \|P_{\lambda,E_T}F_{\beta\alpha}\|^{2N}
  }{N!}
\\
  &=
  \left|
    \mathcal M^{\mathfrak F',\mathfrak F}_{\beta\alpha}
  \right|^2
  \exp\!\left\{
    -\|P_{E_T,M}F_{\beta\alpha}\|^2
  \right\}.
\end{aligned}
\label{eq:dressed-lsz-inclusive-soft-projection}
\end{equation}
Consequently, in the leading logarithmic approximation,
\begin{equation}
  P_{\beta\alpha}
  (\mathfrak F',\mathfrak F;E_T,M)
  =
  \left|
    \mathcal M^{\mathfrak F',\mathfrak F}_{\beta\alpha}
  \right|^2
  \left(\frac{E_T}{M}\right)^{A_{\beta\alpha}}
  \times
  F_{\beta\alpha}^{\rm finite}(E_T/M),
  \label{eq:dressed-lsz-flux-resolved-inclusive-rate}
\end{equation}
where \(F_{\beta\alpha}^{\rm finite}\) has a finite infrared limit and equals
\(1\) at leading logarithmic order.
\end{proposition}

\begin{proof}
At fixed \(\lambda>0\), the soft one-photon space decomposes orthogonally as
\[
  \mathfrak h_{\lambda,M}
  =
  \mathfrak h_{\lambda,E_T}
  \oplus
  \mathfrak h_{E_T,M}.
\]
The coherent soft vector therefore factorizes as the tensor product of a
coherent vector in the unresolved window and a coherent vector in the
resolved window.  The probability of exactly \(N\) unresolved photons and no
resolved photons is
\[
  \exp\{-\|P_{\lambda,E_T}F\|^2\}
  \frac{\|P_{\lambda,E_T}F\|^{2N}}{N!}
  \exp\{-\|P_{E_T,M}F\|^2\}.
\]
Summing over \(N\) gives the exponential series
\[
  \sum_{N=0}^{\infty}
  \exp\{-\|P_{\lambda,E_T}F\|^2\}
  \frac{\|P_{\lambda,E_T}F\|^{2N}}{N!}
  =
  1,
\]
and leaves only the resolved-window no-emission factor
\(\exp\{-\|P_{E_T,M}F\|^2\}\).  Equivalently, writing
\(\|F\|_{\lambda,M}^2=
\|P_{\lambda,E_T}F\|^2+\|P_{E_T,M}F\|^2\) gives the first line of
\eqref{eq:dressed-lsz-inclusive-soft-projection} directly.  The leading
radial integral then gives
\(\exp[-A_{\beta\alpha}\log(M/E_T)]=(E_T/M)^{A_{\beta\alpha}}\), while
nonlogarithmic soft and detector terms are finite matching factors.
\end{proof}

If the sum over \(N\) in
\eqref{eq:dressed-lsz-inclusive-soft-projection} is replaced by one fixed
photon number, the factor
\[
  \exp\{-\|F_{\beta\alpha}\|_{\lambda,M}^2\}
  \frac{\|P_{\lambda,E_T}F_{\beta\alpha}\|^{2N}}{N!}
\]
has no nonzero \(\lambda\downarrow0\) limit when
\(A_{\beta\alpha}>0\).  Fixed photon number is therefore not the physical
charged \(S\)-matrix observable.  The regulator-free object is obtained only
after the detector-degenerate soft states have been summed, precisely as in
Section~\ref{sec:volume-ii-inclusive-probabilities}.

Finally suppose that the detector does not resolve the exact angular-flux
character, but only a finite or measurable cell
\(\mathcal C\) in the spectrum of \(\Phi_\infty\).  Since distinct sharp
flux sectors are orthogonal superselection sectors, the coarse-grained
probability is an incoherent sum or direct integral of
\eqref{eq:dressed-lsz-flux-resolved-inclusive-rate}, not the modulus square
of a sum of amplitudes over different \(\mathfrak F\)'s.  For a finite
cell decomposition the leading-soft formula is
\[
  P_{\beta\alpha}(\mathcal C_{\rm out},\mathcal C_{\rm in};E_T,M)
  =
  \sum_{\mathfrak F\in\mathcal C_{\rm in}}
  \sum_{\mathfrak F'\in\mathcal C_{\rm out}}
  \left|
    \mathcal M^{\mathfrak F',\mathfrak F}_{\beta\alpha}
  \right|^2
  \left(\frac{E_T}{M}\right)^{A_{\beta\alpha}}
  F_{\beta\alpha}^{\rm finite}
\]
with the terms restricted by
\eqref{eq:dressed-lsz-flux-sector-selection}.  Thus detector inclusiveness
does not license replacing angular flux by total charge.  It says instead
that unresolved photons are summed inside the detector resolution, while
different sharp flux sectors contribute as orthogonal alternatives unless the
apparatus prepares and measures a coherent sector-changing observable.

\begin{controlledapproximation}[Flux-cell detector response and charged hard coefficients]
\label{ca:charged-flux-cell-detector-response}
The last step in a charged scattering prediction is not the LSZ residue
itself but the calibrated detector response to the flux-resolved probability.
Fix a finite set \(S\) of sharp incoming/outgoing angular-flux alternatives
which survive the selection rule
\eqref{eq:dressed-lsz-flux-sector-selection}, and write
\[
  P_s
  =
  \left|\mathcal M_s\right|^2
  \left(\frac{E_T}{M}\right)^{A_s}
  F_s^{\rm finite},
  \qquad s\in S,
\]
for the inclusive probability attached to the sharp alternative \(s\).  A
finite-resolution apparatus reports detector cells \(c\in C\).  Its calibrated
flux response is a stochastic matrix \(\mathsf R_{cs}\ge0\) with
\(\sum_c\mathsf R_{cs}=1\).  After ordinary detector backgrounds \(B_c\) and a
declared finite-window remainder \(r_c\), the measured rates have the form
\begin{equation}
  Y_c
  =
  \sum_{s\in S}\mathsf R_{cs}P_s
  +
  B_c
  +
  r_c .
  \label{eq:charged-flux-detector-response-map}
\end{equation}
Thus detector coarse-graining is a positive linear map on probabilities, not
a change of the amplitudes \(\mathcal M_s\).  If a calibrated left inverse
\(\mathsf L\) is available on the retained sector space,
\(\sum_c\mathsf L_{sc}\mathsf R_{cs'}=\delta_{ss'}\), then the reconstructed
sharp probabilities obey
\begin{equation}
  \widehat P_s
  =
  \sum_c\mathsf L_{sc}(Y_c-B_c),
  \qquad
  |\widehat P_s-P_s|
  \le
  \sum_c|\mathsf L_{sc}|\,|r_c|.
  \label{eq:charged-flux-detector-inversion-bound}
\end{equation}
The hard coefficient is then inferred only after the soft and detector
normalizations are divided out:
\begin{equation}
  \left|\widehat{\mathcal M}_s\right|^2
  =
  \frac{\widehat P_s}
       {(E_T/M)^{A_s}F_s^{\rm finite}} .
  \label{eq:charged-flux-hard-coefficient-from-detector}
\end{equation}
If two sharp angular-flux alternatives have identical detector-response
columns, no detector observable in this resolution class separates them; only
their sum in the corresponding response direction is determined.  This is
the finite-resolution version of the earlier flux-sector warning.  A detector
which sees only total charge collapses all alternatives with the same
zero-mode charge into the same column and therefore cannot reconstruct the
angular-flux-resolved coefficient.  Replacing
\eqref{eq:charged-flux-detector-response-map} by
\(|\sum_s\mathcal M_s|^2\) is a different observable, requiring a coherent
sector-changing preparation and readout rather than ordinary flux-cell
detection.
\end{controlledapproximation}

\begin{openproblem}[Haag--Ruelle and LSZ for nonconfining dressed charged sectors]
\label{op:haag-ruelle-lsz-dressed-charged-sectors}
Construct a large-time scattering theory for nonconfining gauge-theory
charged sectors using noncompact gauge-invariant charged creators.  The
construction should prove the analogue of Haag--Ruelle limits for the chosen
dressing class, identify the asymptotic flux labels, and prove a
dressed-correlator reduction formula whose external factors are the residues
or threshold coefficients of the corresponding dressed two-point functions.
It should also identify the detector-inclusive probabilities obtained from
those flux-resolved coefficients after unresolved soft photons and
detector-coarse flux cells have been summed.
The hypothesis ``nonconfining'' means that the charged sector is a genuine
finite-energy physical sector with asymptotic charged data; it excludes
colored charges in a confining phase, for which the physical scattering
problem is the neutral hadron or glueball scattering problem.  In massless QED
this problem must be combined with the infraparticle flux-sector analysis of
Chapter~\ref{ch:volume-ii-infrared-inclusive-qed}.
\end{openproblem}

\section{Wave Operators, Scattering Operator, and Completeness}

Let
\[
  \Omega_{\mathrm{in}},\Omega_{\mathrm{out}}:
  \mathcal F_s(\Hilb_1)\longrightarrow\Hilb
\]
be the Haag--Ruelle wave operators, in the terminology introduced in
Chapter~\ref{ch:haag-ruelle}, produced by Haag--Ruelle limits.  They are
isometric maps.  Their construction has two equivalent topological
descriptions.  First, for each finite velocity-separated Fock vector
\[
  F=\psi_1\odot\cdots\odot\psi_n
\]
represented by regular creators \(B_j\) and packets \(h_j\), the vector family
\[
  \Omega_{{\rm out},t}^{\rm HR}F
  =
  B_{1,t}(h_1)\cdots B_{n,t}(h_n)\vac
\]
converges in the Hilbert norm of \(\Hilb\).  Second, after these vector limits
are packaged as maps from the algebraic velocity-separated Fock domain to
\(\Hilb\), the same assertion is strong convergence of the approximating maps:
\begin{equation}
  \Omega_{{\rm out},t}^{\rm HR}F
  \longrightarrow
  \Omega_{\rm out}F
  \quad\text{in }\Hilb
  \quad\text{for every fixed }F.
  \label{eq:algebraic-haag-ruelle-strong-operator-limit}
\end{equation}
The incoming map is obtained by the same statement with \(t\to-\infty\).
Weak convergence of matrix elements follows from this norm convergence for
each \(F\), while weak convergence alone would not determine the isometric
identification.  No operator-norm limit over the unit sphere of
\(\mathcal F_s(\Hilb_1)\) is part of the theorem.  Their ranges
\[
  \Hilb_{\mathrm{in}}=\operatorname{Ran}\Omega_{\mathrm{in}},
  \qquad
  \Hilb_{\mathrm{out}}=\operatorname{Ran}\Omega_{\mathrm{out}}
\]
are the incoming and outgoing scattering subspaces.  The comparison operator
\[
  S=\Omega_{\mathrm{out}}^*\Omega_{\mathrm{in}}
  \quad\text{on}\quad \mathcal F_s(\Hilb_1)
\]
is unitary on the sector when \(\Hilb_{\mathrm{in}}=\Hilb_{\mathrm{out}}\).  Its
wave-packet matrix elements are
\[
\begin{aligned}
&\left\langle
  \varphi_1\odot\cdots\odot\varphi_m,\,
  S(\psi_1\odot\cdots\odot\psi_n)
  \right\rangle_{\mathcal F_s(\Hilb_1)}
\\
&\qquad =
  \left\langle
  \Omega_{\mathrm{out}}(\varphi_1\odot\cdots\odot\varphi_m),\,
  \Omega_{\mathrm{in}}(\psi_1\odot\cdots\odot\psi_n)
  \right\rangle_{\Hilb}.
\end{aligned}
\]
This is the nonperturbative definition of the scattering matrix element in
the massive isolated-shell sector.

Asymptotic completeness is a further assertion:
\[
  \Hilb_{\rm sector}=\Hilb_{\mathrm{in}}=\Hilb_{\mathrm{out}}.
\]
It is a dynamical property of the model or sector.  For nontrivial interacting
massive local Wightman theories in four spacetime dimensions, full asymptotic
completeness is presently an open problem rather than a theorem following from
the axioms.  Existing proofs are confined to restricted model classes, notably
constructive or integrable models in lower spacetime dimensions and related
weakened-locality settings.  Bound states require additional isolated shells.
Confinement can remove charged or colored one-particle shells from the
observable spectrum while retaining hadronic shells.  Long-range gauge forces
and infraparticle spectra require modified asymptotic state spaces or detector
observables; the charged-QED and physical-QCD versions are the open
asymptotic-completeness problems formulated in
Open Problems~\ref{op:qed-charged-asymptotic-completeness}
and~\ref{op:qcd-asymptotic-completeness}.  Each case changes the large-time
construction rather than the meaning of the massive theorem above.

\section{Interface With LSZ and Perturbation Theory}

The Haag--Ruelle construction defines the Hilbert-space target of LSZ.  If a
local field or local observable has a nonzero vacuum-to-one-particle matrix
element on the isolated shell, then its two-point function contains the
corresponding one-particle pole or spectral atom.  LSZ extracts those external
singularities from time-ordered correlation functions and thereby computes
matrix elements of the scattering operator already defined by wave operators.

The logical dependencies are
\[
\begin{gathered}
  \text{local net and spectrum}
  \longrightarrow
  \text{Haag--Ruelle wave operators}
  \longrightarrow S,
  \\[0.35em]
  \text{time-ordered Green functions}
  \xrightarrow{\ \text{LSZ external residues}\ }
  \langle{\mathrm{out}}|S|{\mathrm{in}}\rangle,
  \\[0.35em]
  \text{renormalized perturbation theory}
  \longrightarrow
  \text{Green-function expansion}
  \longrightarrow
  \text{expansion of } \langle{\mathrm{out}}|S|{\mathrm{in}}\rangle .
\end{gathered}
\]
Thus Feynman graphs for Green functions acquire a scattering interpretation
only after the asymptotic Hilbert-space data and the reduction theorem are in
place.  The same order applies whether the theory is presented by a
Lagrangian, a Euclidean construction satisfying reconstruction hypotheses, a
local net, or another nonperturbative definition.
