\chapter{Superselection Sectors and Locality Properties}
\label{ch:volume-iv-superselection-sectors-locality-properties}
\ConventionsLorentzian


\section{Representations as Sectors}

Let \(M=\mathbb M^D\) be Minkowski spacetime, let
\(\mathcal O\mapsto\Obs(\mathcal O)\) be a local observable net, and let
\(\Obs_{\rm qloc}\) be its quasilocal algebra.  Let
\[
  \pi_0:\Obs_{\rm qloc}\to \mathcal B(\Hilb_0)
\]
be a vacuum representation.  A representation
\[
  \pi:\Obs_{\rm qloc}\to \mathcal B(\Hilb_\pi)
\]
defines a representation sector.  Two representations belong to the same
sector when they are unitarily equivalent:
\[
  \pi_1\simeq\pi_2
  \quad\Longleftrightarrow\quad
  \exists\,V:\Hilb_{\pi_1}\to\Hilb_{\pi_2}
  \text{ unitary with }V\pi_1(A)=\pi_2(A)V .
\]
For irreducible representations, inequivalent sectors have no nonzero
observable intertwiner between them.  More generally, disjoint
representations have no unitarily equivalent nonzero subrepresentations; this
is the representation-theoretic content of superselection by the observable
algebra.

In massive relativistic theories, the sectors relevant for ordinary charges
are often captured by the Doplicher--Haag--Roberts localization criterion.  A
representation \(\pi\) is localized in a bounded region \(\mathcal O\) relative to the
vacuum if
\begin{equation}
  \pi|_{\Obs(\mathcal O^\perp)}
  \simeq
  \pi_0|_{\Obs(\mathcal O^\perp)} ,
  \label{eq:dhr-localization}
\end{equation}
where \(\mathcal O^\perp\) denotes the spacelike complement and
\(\Obs(\mathcal O^\perp)\) is the algebra generated by local algebras
\(\Obs(\mathcal U)\) with \(\mathcal U\subset\mathcal O^\perp\).  It is
transportable if an equivalent representation can be localized in any other
suitable bounded region.  A transportable localized sector is a charge class
whose localization region may be changed without changing its sector.

Figure~\ref{fig:dhr-localized-sector} records the geometric content of the DHR
localization condition in a \(1+1\)-dimensional slice.  The relevant exterior
region is the spacelike complement \(\mathcal O^\perp\), not the set-theoretic
complement of \(\mathcal O\): observables supported there have no causal
contact with the bounded localization region, and the sector representation
restricts to the vacuum representation on the algebra generated by those
observables.

\begin{figure}[H]
  \centering
  \resizebox{0.86\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      arrow/.style={-{Latex[length=2mm]},line width=0.85pt},
      lab/.style={fill=white,inner sep=1.6pt,align=center}
    }
    \draw[->] (-4.7,0)--(4.7,0) node[right] {\(x^1\)};
    \draw[->] (0,-2.35)--(0,2.32);
    \node[anchor=west] at (0.12,2.30) {\(x^0\)};

    \fill[orange!8] (-4.65,-2.10)--(-2.55,-2.10)--(-0.55,0)
      --(-2.55,2.10)--(-4.65,2.10)--cycle;
    \fill[orange!8] (4.65,-2.10)--(2.55,-2.10)--(0.55,0)
      --(2.55,2.10)--(4.65,2.10)--cycle;
    \node[lab,text=orange!80!black] at (-3.20,1.28)
      {\(\mathcal O^\perp\)\\spacelike complement};
    \node[lab,text=orange!80!black] at (3.20,1.28)
      {\(\mathcal O^\perp\)\\same algebraic region};

    \fill[gray!10] (0,2.10)--(2.55,0)--(0,-2.10)--(-2.55,0)--cycle;
    \draw[gray!55,line width=0.8pt,dashed] (0,2.10)--(2.55,0)--(0,-2.10)--(-2.55,0)--cycle;
    \node[lab,text=gray!65!black] at (0,-1.52)
      {causal shadow of \(\mathcal O\)};

    \fill[blue!10] (0,0.70)--(0.78,0)--(0,-0.70)--(-0.78,0)--cycle;
    \draw[blue!65!black,line width=1.1pt] (0,0.70)--(0.78,0)--(0,-0.70)--(-0.78,0)--cycle;
    \node[lab,text=blue!65!black] at (0,0.98)
      {localized charge\\in \(\mathcal O\)};

    \node[lab] (eq) at (0,3.10)
      {\(\pi|_{\Obs(\mathcal O^\perp)}
        \simeq
        \pi_0|_{\Obs(\mathcal O^\perp)}\)};
    \draw[arrow,orange!80!black] (-3.15,1.65) .. controls (-2.25,2.80) and (-1.45,3.12) .. (eq.west);
    \draw[arrow,orange!80!black] (3.15,1.65) .. controls (2.25,2.80) and (1.45,3.12) .. (eq.east);
  \end{tikzpicture}
  }
  \caption{DHR localization in a spacetime slice.  The sector may differ from
  the vacuum inside a bounded localization region \(\mathcal O\), while on the
  algebra generated by observables in the spacelike complement
  \(\mathcal O^\perp\) its representation is unitarily equivalent to the vacuum
  representation.}
  \label{fig:dhr-localized-sector}
\end{figure}

\section{Endomorphisms, Fusion, and Intertwiners}

Assume Haag duality in the vacuum representation,
\[
  \mathcal R(\mathcal O)=\mathcal R(\mathcal O^\perp)' ,
\]
and assume the representation \(\pi\) satisfies the localization and
transportability condition above.  After identifying
\(\pi|_{\Obs(\mathcal O^\perp)}\) with
\(\pi_0|_{\Obs(\mathcal O^\perp)}\), and after replacing
\(\Obs_{\rm qloc}\) by its image in the vacuum representation, the action of
\(\pi\) may be represented by an endomorphism
\[
  \rho:\Obs_{\rm qloc}\to\Obs_{\rm qloc}
\]
localized in \(\mathcal O\), meaning that
\[
  \rho(A)=A,\qquad A\in\Obs(\mathcal O^\perp).
\]
The reason Haag duality appears in this sentence is the containment argument
of Chapter~\ref{ch:volume-iv-aqft-local-nets-states}.  If
\(\mathcal O_0\) localizes \(\rho\) and
\(\mathcal O_0\subset\mathcal O_1\), then for
\(A\in\mathcal R(\mathcal O_1)\) and
\(B\in\mathcal R(\mathcal O_1^\perp)\),
\[
  \rho(A)B
  =
  \rho(A)\rho(B)
  =
  \rho(AB)
  =
  \rho(BA)
  =
  B\rho(A).
\]
Hence \(\rho(A)\in\mathcal R(\mathcal O_1^\perp)'\).  Haag duality turns this
exterior-commutation conclusion into
\(\rho(A)\in\mathcal R(\mathcal O_1)\).  Similarly, an intertwiner
\(T\rho(\,\cdot\,)=\sigma(\,\cdot\,)T\) between endomorphisms localized in
\(\mathcal O\) commutes with \(\mathcal R(\mathcal O^\perp)\), and Haag
duality places \(T\) in \(\mathcal R(\mathcal O)\).  Without this step the
same construction naturally lands in the dual net
\(\mathcal R^d(\mathcal O)=\mathcal R(\mathcal O^\perp)'\).
Composition of localized endomorphisms defines fusion:
\[
  \rho_1\otimes\rho_2=\rho_1\circ\rho_2 .
\]
We use ordinary function-composition order:
\[
  (\rho_1\circ\rho_2)(A)=\rho_1(\rho_2(A)),
  \qquad A\in\Obs_{\rm qloc}.
\]
Thus \(\rho_2\) acts first and \(\rho_1\) acts second.
An intertwiner from \(\rho\) to \(\sigma\) is an observable
\(T\in\Obs_{\rm qloc}\) such
that
\[
  T\rho(A)=\sigma(A)T,\qquad A\in\Obs_{\rm qloc}.
\]
The intertwiners are morphisms between charge sectors.  With direct sums,
subobjects, conjugates, and a tensor product, transportable localized
endomorphisms form a tensor category.  Locality gives the exchange
isomorphism
\[
  \varepsilon_{\rho,\sigma}:\rho\circ\sigma\to\sigma\circ\rho
\]
when the localization regions are spacelike separated and then transported
back to common reference regions.  In spacetime dimension at least three, this
exchange is symmetric when the DHR hypotheses used above are supplemented by
transportability, Haag duality for the vacuum net, finite statistics, and the
path-connectedness of the relevant spacelike localization configurations.
The last condition is the topological input that makes the two exchanges
\(\rho\) around \(\sigma\) homotopic.  In two dimensions and for
lower-codimension extended sectors, the corresponding configuration spaces
need not have this symmetry; the exchange maps are then represented by braid
group actions rather than by the symmetric group.

\section{Statistics Operators and Conjugates}

The exchange isomorphism contains the statistics of the sector.  For a
localized endomorphism \(\rho\), choose a transported copy
\(\rho'\) localized spacelike to \(\rho\).  Locality gives an intertwiner
\[
  \varepsilon_{\rho,\rho}:\rho\circ\rho\to\rho\circ\rho
\]
after transporting \(\rho'\) back to the reference region.  In spacetime
dimension at least three, the DHR exchange operators satisfy symmetric tensor
category identities, so irreducible sectors have Bose, Fermi, or more general
parastatistics encoded by finite-dimensional permutation-group
representations.

For objects \(\eta,\theta\), the morphism space
\(\operatorname{Hom}(\eta,\theta)\) is the space of intertwiners from
\(\eta\) to \(\theta\).  The symbol \(1_\eta\) denotes the identity morphism in
\(\operatorname{Hom}(\eta,\eta)\); in the endomorphism realization it is the
unit observable viewed as an intertwiner \(\eta\to\eta\).  A sector \(\rho\)
has a conjugate \(\bar\rho\) when there are intertwiners
\[
  R:\operatorname{id}\to \bar\rho\circ\rho,
  \qquad
  \bar R:\operatorname{id}\to \rho\circ\bar\rho
\]
satisfying the conjugate equations
\[
  \bar R^*\rho(R)=1_\rho,
  \qquad
  R^*\bar\rho(\bar R)=1_{\bar\rho}.
\]
The minimal value of \(\|R\|\,\|\bar R\|\) defines the statistical dimension
\[
  d(\rho).
\]
In the Doplicher--Haag--Roberts setting, saying that a sector has finite
statistics is shorthand for a categorical finiteness package, not only for a
choice of exchange sign.  In the form used below it means that the sector lies
in a \(C^*\)-tensor category with finite-dimensional intertwiner spaces, admits
a conjugate satisfying the conjugate equations, and has finite statistical
dimension \(d(\rho)<\infty\).  For reducible sectors, the statement is applied
after decomposing into irreducible subobjects and requiring only finitely many
irreducible summands in each object under consideration.
For sectors reconstructed from a compact internal group, \(d(\rho)\) is the
ordinary dimension of the corresponding irreducible group representation.
Thus charge conjugation, fusion multiplicities, and statistics are encoded in
the tensor category of localized endomorphisms before any charged field
coordinates are chosen.

\section{Observable Algebra, Field Algebra, and Compact Internal Group}

The observable algebra \(\Obs_{\rm qloc}\) contains neutral operators.  Charged fields,
when they are introduced as operators, can be organized in an enlarged algebra
\[
  \mathcal F
\]
with a compact internal group \(G\) acting by automorphisms.  The observable
algebra is the fixed-point algebra
\[
  \Obs_{\rm qloc}=\mathcal F^G .
\]

\begin{quotedtheorem}[Doplicher--Roberts reconstruction]
\label{thm:doplicher-roberts-reconstruction}
\emph{Proof references.}\footnote{The DHR sector analysis is developed in
S.~Doplicher, R.~Haag, and J.~E.~Roberts, ``Local observables and particle
statistics I,'' \emph{Communications in Mathematical Physics} \textbf{23}
(1971), 199--230, and ``Local observables and particle statistics II,''
\emph{Communications in Mathematical Physics} \textbf{35} (1974), 49--85.
The categorical compact-group duality is proved in S.~Doplicher and
J.~E.~Roberts, ``A new duality theory for compact groups,''
\emph{Inventiones Mathematicae} \textbf{98} (1989), 157--218.  The
field-algebra reconstruction in the AQFT setting is proved in S.~Doplicher
and J.~E.~Roberts, ``Why there is a field algebra with a compact gauge group
describing the superselection structure in particle physics,''
\emph{Communications in Mathematical Physics} \textbf{131} (1990), 51--107
\cite{DHR1971,DHR1974,DR1989,DR1990}.}
Let \(D\ge3\).  Let the vacuum representation of the observable net be
irreducible and satisfy Haag duality for double cones.  Let \(\mathcal C\) be
the category of DHR sectors represented by transportable localized
endomorphisms of \(\Obs_{\rm qloc}\).  Equivalently, their associated
representations satisfy the localization criterion
\eqref{eq:dhr-localization} in some bounded double cone and are transportable
to every other such double cone; in the endomorphism realization they act
trivially on the algebra of the spacelike complement.  Assume that
\(\mathcal C\) has the following structure:
\begin{enumerate}[label=\textup{(\roman*)}]
\item it is a symmetric \(C^*\)-tensor category with simple tensor unit;
\item each intertwiner space \(\operatorname{Hom}(\rho,\sigma)\) is
finite-dimensional;
\item the category is closed under finite direct sums and subobjects;
\item every irreducible object has a conjugate object satisfying the conjugate
equations and has finite statistical dimension \(d(\rho)<\infty\);
\item every object of \(\mathcal C\) is a finite direct sum of irreducible
subobjects of the preceding kind.
\end{enumerate}
Then there is a compact group
\[
  G=\operatorname{Aut}^{\otimes,*}(\omega_{\rm fib})
\]
for a faithful symmetric tensor \(*\)-functor
\(\omega_{\rm fib}:\mathcal C\to{\rm Hilb}_{\rm fd}\), where
\({\rm Hilb}_{\rm fd}\) denotes finite-dimensional Hilbert spaces, and
\[
  \mathcal C\simeq {\rm Rep}_{\rm fd}(G)
\]
as symmetric \(C^*\)-tensor categories.  Moreover, after the standard
Doplicher--Roberts field-algebra reconstruction, there is a \(C^*\)-algebra
\(\mathcal F\) carrying a continuous action of \(G\) by automorphisms such that
\[
  \Obs_{\rm qloc}\simeq \mathcal F^G .
\]
The irreducible DHR sectors correspond to irreducible finite-dimensional
unitary representations of \(G\), and charged field multiplets in
\(\mathcal F\) transform according to these representations.
\end{quotedtheorem}

In this quoted theorem, the word gauge means a reconstruction of charged field
coordinates from the neutral observable category; physical observables remain
the \(G\)-invariant elements.  The compact group description is therefore not
an input at this stage.  It is a conclusion from locality, transportability,
Haag duality, symmetric exchange, and finite statistics.

\paragraph{Reconstruction logic and theorem boundary.}
The reconstruction has two logically separate components.  The first is the
QFT-specific passage from localized representations to a rigid symmetric
\(C^*\)-tensor category.  Haag duality is the step that permits a DHR
representation localized in \(\mathcal O\) to be represented by an
endomorphism \(\rho\) of the vacuum quasilocal algebra acting trivially on
\(\Obs(\mathcal O^\perp)\).  Transportability makes the localization region a
choice of representative rather than part of the sector.  Locality and the
path-connectedness of spacelike configurations in \(D\ge3\) give a symmetric
exchange isomorphism, while finite statistics supplies conjugates and finite
statistical dimension.  These are the hypotheses that must be verified in a
model before the theorem applies.

The second component is categorical duality.  Once the category
\(\mathcal C\) has the properties listed in
Theorem~\ref{thm:doplicher-roberts-reconstruction}, the compact group
\[
  G=\operatorname{Aut}^{\otimes,*}(\omega_{\rm fib})
\]
is obtained as the group of unitary tensor automorphisms of a faithful fiber
functor \(\omega_{\rm fib}\).  Its topology is the compact topology inherited
from the product of the unitary groups
\(\prod_{\rho\in{\rm Obj}(\mathcal C)}U(\omega_{\rm fib}(\rho))\), with the
closed conditions expressing naturality and tensor compatibility.  The
equivalence \(\mathcal C\simeq{\rm Rep}_{\rm fd}(G)\) is therefore a
Tannakian output of the observable sector category.

The field algebra reconstructed from \(\mathcal C\) can be recognized
locally by its charged multiplets.  For an irreducible sector \(\rho\) with
statistical dimension \(d(\rho)\), which is the dimension of the corresponding
irreducible representation after the Doplicher--Roberts equivalence, the
reconstructed algebra contains a finite-dimensional Hilbert space of charged
operators
\[
  \mathcal H_\rho^{\rm ch}
  =
  \operatorname{span}\{\psi_{\rho,1},\ldots,\psi_{\rho,d(\rho)}\}
  \subset\mathcal F
\]
whose elements satisfy the intertwining relation
\begin{equation}
  \rho(A)\psi_{\rho,i}=\psi_{\rho,i}A,
  \qquad A\in\Obs_{\rm qloc}.
  \label{eq:dr-charged-field-intertwining}
\end{equation}
After choosing an orthonormal multiplet, one has the Cuntz-type normalization
\begin{equation}
  \psi_{\rho,i}^*\psi_{\rho,j}=\delta_{ij}\mathbf 1,
  \qquad
  \sum_{i=1}^{d(\rho)}\psi_{\rho,i}\psi_{\rho,i}^*=\mathbf 1,
  \label{eq:dr-charged-field-cuntz-relations}
\end{equation}
in the finite-statistics case, and the observable endomorphism is recovered
from the charged fields by
\begin{equation}
  \rho(A)=\sum_{i=1}^{d(\rho)}\psi_{\rho,i}A\psi_{\rho,i}^* .
  \label{eq:dr-endomorphism-from-charged-fields}
\end{equation}
The \(G\)-action rotates the multiplet by the representation corresponding to
\(\rho\):
\begin{equation}
  \alpha_g(\psi_{\rho,i})
  =
  \sum_{j=1}^{d(\rho)}
  \psi_{\rho,j}\,D_\rho(g)_{ji},
  \qquad
  g\in G .
  \label{eq:dr-compact-group-acts-on-charged-multiplet}
\end{equation}
Equations
\eqref{eq:dr-charged-field-intertwining}--\eqref{eq:dr-compact-group-acts-on-charged-multiplet}
are not additional assumptions on observables; they are the operator form of
the reconstructed charged coordinates.  The observable algebra remains the
fixed algebra \(\mathcal F^G\), and the physical sector information is already
present in \(\mathcal C\).

\paragraph{Local charged-field core.}
The preceding formulae become a local field net only after the localization
regions of the DHR endomorphisms are used.  Fix a double cone \(\mathcal O\)
and choose, for each irreducible DHR sector \(\rho\) localizable in
\(\mathcal O\), a charged multiplet
\[
  \mathcal H_\rho^{\rm ch}(\mathcal O)
  =
  \operatorname{span}\{\psi_{\rho,1}^{\mathcal O},\ldots,
  \psi_{\rho,d(\rho)}^{\mathcal O}\}
\]
satisfying \eqref{eq:dr-charged-field-intertwining} with \(\rho\) localized
in \(\mathcal O\).  The algebraic local core is the finite linear span
\begin{equation}
  \mathcal F_{\rm alg}(\mathcal O)
  =
  \operatorname{span}\{
    A\psi_{\rho,i}^{\mathcal O}:
    A\in\mathcal R(\mathcal O),\
    \rho\in{\rm Irr}(\mathcal C),\
    1\le i\le d(\rho)
  \}.
  \label{eq:dr-local-field-core}
\end{equation}
This is a core, not yet a completed local \(C^*\)-algebra.  Its multiplication
is forced by the intertwining relation.  If \(A,B\in\mathcal R(\mathcal O)\),
then
\begin{equation}
  (A\psi_{\rho,i}^{\mathcal O})(B\psi_{\sigma,j}^{\mathcal O})
  =
  A\,\rho(B)\,
  \psi_{\rho,i}^{\mathcal O}\psi_{\sigma,j}^{\mathcal O}.
  \label{eq:dr-local-field-core-product}
\end{equation}
Haag duality and the containment calculation above are used here in a
load-bearing way: since \(\rho\) is localized in \(\mathcal O\),
\(\rho(B)\in\mathcal R(\mathcal O)\).  The product
\(\psi_{\rho,i}^{\mathcal O}\psi_{\sigma,j}^{\mathcal O}\) transforms in the
sector \(\rho\circ\sigma\), because
\[
  (\rho\circ\sigma)(A)
  \psi_{\rho,i}^{\mathcal O}\psi_{\sigma,j}^{\mathcal O}
  =
  \psi_{\rho,i}^{\mathcal O}\sigma(A)\psi_{\sigma,j}^{\mathcal O}
  =
  \psi_{\rho,i}^{\mathcal O}\psi_{\sigma,j}^{\mathcal O}A .
\]
After decomposing \(\rho\circ\sigma\) into irreducible sectors, the
Clebsch--Gordan intertwiners in
\(\operatorname{Hom}(\tau,\rho\circ\sigma)\) rewrite the right side of
\eqref{eq:dr-local-field-core-product} as a finite sum of generators of the
form \eqref{eq:dr-local-field-core}.  Thus fusion in the DHR category is the
multiplication law of local charged coordinates.

The involution uses the conjugate sector.  Finite statistics supplies
standard conjugate data
\[
  R:\operatorname{id}\to \bar\rho\circ\rho,
  \qquad
  \bar R:\operatorname{id}\to \rho\circ\bar\rho ,
\]
and the adjoint of a \(\rho\)-charged generator is expanded, through this
conjugate data and local intertwiners, as a finite linear combination of
\(\bar\rho\)-charged generators with coefficients in
\(\mathcal R(\mathcal O)\).  Finite statistics therefore supplies the
adjoint-closing structure required by the field-algebra theorem, in addition
to the exchange data.

The reconstructed compact group acts on the local core by
\[
  \alpha_g(A)=A,\qquad
  \alpha_g(\psi_{\rho,i}^{\mathcal O})
  =
  \sum_j\psi_{\rho,j}^{\mathcal O}D_\rho(g)_{ji}.
\]
The Haar conditional expectation on the local core is
\begin{equation}
  E_{\mathcal O}(F)
  =
  \int_G\alpha_g(F)\,\dd g.
  \label{eq:dr-local-haar-expectation}
\end{equation}
By Schur orthogonality it kills every nontrivial irreducible charged block and
keeps the trivial block.  Consequently
\[
  E_{\mathcal O}(\mathcal F_{\rm alg}(\mathcal O))
  =
  \mathcal R(\mathcal O),
  \qquad
  \mathcal F_{\rm alg}(\mathcal O)^G=\mathcal R(\mathcal O)
\]
at the algebraic core level.  The Doplicher--Roberts field-algebra theorem is
the statement that this core can be completed, with compatible choices as
\(\mathcal O\) varies, to a local charged-field net whose \(G\)-fixed subnet
is the original observable net.  The nontrivial analytic work is not the
formal multiplication above; it is the proof that the transported charged
multiplets, exchange operators, adjoints, \(C^*\)-norm, and locality relations
can be chosen coherently across all regions.

\paragraph{Charged locality and observable locality.}
For spacelike separated double cones, the charged field generators need not
commute as ordinary neutral observables.  Their exchange is controlled by the
DHR symmetry
\[
  \varepsilon_{\rho,\sigma}:\rho\circ\sigma\to\sigma\circ\rho .
\]
In dimension \(D\ge3\), under the symmetric finite-statistics hypotheses of
Theorem~\ref{thm:doplicher-roberts-reconstruction}, this exchange reduces to
the Bose/Fermi/parastatistical symmetry encoded by the reconstructed compact
group and its central grading.  The invariant subnet is local because the
exchange factors cancel in the trivial representation channel.  Thus the
locality statement for observables is recovered as the invariant part of the
charged-field exchange mechanism.

\paragraph{Neutral-channel locality in the reconstructed core.}
The preceding cancellation is a categorical statement before it is an
operator-closure theorem.  Let \(V_\rho\) and \(V_\sigma\) be the finite
charged multiplet spaces carrying the compact-group representations
\(D_\rho\) and \(D_\sigma\).  The projection onto the neutral component of the
product multiplet is
\begin{equation}
  P_{\mathbf1}^{\rho,\sigma}
  =
  \int_G D_\rho(g)\otimes D_\sigma(g)\,\dd g
  \in \operatorname{End}(V_\rho\otimes V_\sigma).
  \label{eq:dr-neutral-channel-projector}
\end{equation}
For irreducible \(\rho,\sigma\), Schur orthogonality says that
\(P_{\mathbf1}^{\rho,\sigma}=0\) unless
\(\sigma\simeq\bar\rho\); in the conjugate case its image is the
one-dimensional tensor-unit summand
\[
  \operatorname{Hom}_G(\mathbb C,V_\rho\otimes V_{\bar\rho}) .
\]
Applying \(E_{\mathcal O}\) in
\eqref{eq:dr-local-haar-expectation} to a product of charged local
generators is exactly the insertion of these projectors on the charged
indices after the fusion decomposition.  Thus an observable local element is
not an arbitrary product of charged fields; it is the tensor-unit component
of such a product, with all nontrivial charged blocks removed by Haar
projection.

Now take \(F_1\in\mathcal F_{\rm alg}(\mathcal O_1)^G\) and
\(F_2\in\mathcal F_{\rm alg}(\mathcal O_2)^G\) with
\(\mathcal O_1\perp\mathcal O_2\).  The charged-field exchange operation
first transports the charged tensor factors by
\(\varepsilon_{\rho,\sigma}\).  But \(F_1\) and \(F_2\) lie in tensor-unit
subobjects after the projections
\eqref{eq:dr-neutral-channel-projector}.  Naturality of the symmetric
braiding and simplicity of the tensor unit give
\[
  \varepsilon_{\mathbf1,\mathbf1}=\operatorname{id}_{\mathbf1}.
\]
Hence no charge matrix remains on the invariant channel: the categorical
exchange reduces to the ordinary equality required for observable locality.
This proves only the algebraic neutral-channel mechanism in the DR core.  The
hard local-QFT theorem remains the coherent \(C^*\)-completion and regionwise
choice of charged multiplets for which the charged exchange relations and
the invariant local observable net survive completion.

\paragraph{Finite pointed check of the compact group output.}
A useful local model of the categorical part is the following pointed
symmetric \(C^*\)-tensor category.  Fix an integer \(N\geq2\).  Let the simple
objects be \(X_q\), \(q\in\mathbb Z/N\mathbb Z\), with
\[
  X_q\otimes X_r\simeq X_{q+r},
  \qquad
  X_0=\mathbf 1,
  \qquad
  \operatorname{Hom}(X_q,X_r)=
  \begin{cases}
    \mathbb C, & q=r,\\
    0, & q\ne r .
  \end{cases}
\]
Take the fiber functor \(\omega_{\rm fib}(X_q)=\mathbb C\).  A unitary tensor
automorphism \(u\in\operatorname{Aut}^{\otimes,*}(\omega_{\rm fib})\) is then a
choice of phases \(u_q\in U(1)\) satisfying
\[
  u_0=1,
  \qquad
  u_{q+r}=u_q u_r .
\]
Thus \(u_q=u_1^q\) and \(u_1^N=1\).  The reconstructed compact group is
\begin{equation}
  G\simeq \mu_N=\{\zeta\in U(1):\zeta^N=1\},
  \label{eq:dr-pointed-zn-reconstructed-group}
\end{equation}
and \(X_q\) is the one-dimensional representation
\[
  \zeta\longmapsto \zeta^q .
\]
If the corresponding algebraic charged-field core is written as a finite
grading
\begin{equation}
  \mathcal F_{\rm alg}
  =
  \bigoplus_{q\in\mathbb Z/N\mathbb Z}\mathcal F_q,
  \qquad
  \mathcal F_q\mathcal F_r\subset \mathcal F_{q+r},
  \qquad
  \alpha_\zeta(F_q)=\zeta^q F_q,
  \quad F_q\in\mathcal F_q ,
  \label{eq:dr-pointed-zn-field-grading}
\end{equation}
then the fixed algebra is exactly the degree-zero summand:
\begin{equation}
  \mathcal F_{\rm alg}^{\mu_N}=\mathcal F_0 .
  \label{eq:dr-pointed-zn-fixed-algebra}
\end{equation}
Indeed, averaging \(F_q\) over \(G\) multiplies it by
\[
  \frac1N\sum_{\zeta^N=1}\zeta^q
  =
  \delta_{q,0}.
\]
This computation functions as a finite-dimensional diagnostic for the theorem's
output: the compact group is recovered from the tensor compatibility of the
sector fiber functor, and invariants project precisely onto neutral degree.

\paragraph{Finite nonabelian check of the fiber functor.}
The compact-group output also has genuinely nonabelian content.  Take the
finite nonabelian group \(S_3\).  Its irreducible representations are the
trivial representation \(\mathbf 1\), the sign representation \(\epsilon\),
and the two-dimensional standard representation \(V\).  With conjugacy
classes ordered as
\[
  (e),\qquad (12),\qquad (123),
\]
their characters are
\[
\begin{array}{c|ccc}
  & e & (12) & (123)\\ \hline
  \mathbf 1 & 1 & 1 & 1\\
  \epsilon & 1 & -1 & 1\\
  V & 2 & 0 & -1 .
\end{array}
\]
The inner product
\[
  \langle \chi,\psi\rangle_{S_3}
  =
  \frac16\sum_{g\in S_3}\chi(g)\overline{\psi(g)}
\]
gives the tensor-product rules
\begin{equation}
  \epsilon\otimes\epsilon\simeq \mathbf 1,
  \qquad
  \epsilon\otimes V\simeq V,
  \qquad
  V\otimes V\simeq \mathbf 1\oplus\epsilon\oplus V .
  \label{eq:dr-s3-representation-ring}
\end{equation}
For each group element \(g\), evaluation gives a unitary tensor
automorphism of the forgetful fiber functor
\[
  \omega_{\rm fib}:{\rm Rep}_{\rm fd}(S_3)\to{\rm Hilb}_{\rm fd},
  \qquad
  \eta_g(W)=D_W(g).
\]
Tensor compatibility is the identity
\[
  D_{W_1\otimes W_2}(g)=D_{W_1}(g)\otimes D_{W_2}(g),
\]
and naturality says that every intertwiner, in particular the projection
maps realizing \eqref{eq:dr-s3-representation-ring}, commutes with
\(\eta_g\).  The standard representation is faithful, so the six
automorphisms \(\eta_g\) are distinct.  The converse statement, that every
unitary tensor natural automorphism of the forgetful functor is evaluation at
a unique \(g\in S_3\), is the finite Tannakian part of the
Doplicher--Roberts theorem in this example.  The calculation above makes
visible what the theorem reconstructs: a nonabelian group is encoded by the
compatibility of the multiplet transformations with all fusion and
intertwiner maps, not by an externally assigned charge label.

The fixed-point projection is likewise representation-theoretic rather than
degree-theoretic.  If a finite charged-field core decomposes into matrix
coefficient spaces for irreducible \(S_3\)-representations, Haar averaging is
\[
  E_{S_3}(F)
  =
  \frac16\sum_{g\in S_3}\alpha_g(F).
\]
For a nontrivial irreducible representation \(W\), Schur orthogonality gives
\[
  \frac16\sum_{g\in S_3}D_W(g)_{ab}=0,
\]
whereas the trivial representation is fixed.  Thus averaging projects onto
the \(\mathbf1\)-isotypic, observable part.  In the pointed
\(\mathbb Z/N\mathbb Z\) model this projection is the elementary root-of-unity
sum; in the nonabelian \(S_3\) model it is the vanishing of nontrivial matrix
coefficients.

\paragraph{Finite nonabelian charged-coordinate core.}
The preceding \(S_3\) calculation can be made into a finite field-core model,
with the observable algebra suppressed to \(\mathbb C\), by taking
\[
  \mathcal F_{\rm ch}=\operatorname{Fun}(S_3)
\]
with pointwise multiplication.  The reconstructed compact group acts on the
right:
\begin{equation}
  (\alpha_h f)(g)=f(gh),
  \qquad f\in\operatorname{Fun}(S_3),\quad g,h\in S_3 .
  \label{eq:dr-s3-regular-right-action}
\end{equation}
The fixed algebra is the constants, since right translations act transitively
on \(S_3\).  For an irreducible finite-dimensional representation \(W\), its matrix
coefficients
\[
  f^W_{ab}(g)=D_W(g)_{ab}
\]
span a right \(S_3\)-multiplet:
\begin{equation}
  \alpha_h(f^W_{ab})
  =
  \sum_c f^W_{ac}\,D_W(h)_{cb}.
  \label{eq:dr-s3-matrix-coefficient-right-action}
\end{equation}
Thus the column index is the charged multiplet index, while the row index
records the multiplicity with which \(W\) appears in the finite regular
field-core model.  Haar projection is the finite conditional expectation
\begin{equation}
  E_{S_3}(f)
  =
  \frac1{6}\sum_{h\in S_3}\alpha_h(f),
  \qquad
  E_{S_3}(f^W_{ab})=
  \begin{cases}
    1,& W=\mathbf 1,\\
    0,& W\ne\mathbf 1,
  \end{cases}
  \label{eq:dr-s3-regular-haar-expectation}
\end{equation}
with the second line understood after choosing the constant matrix coefficient
for \(W=\mathbf1\).  Multiplication of charged coordinates is the
matrix-coefficient form of tensor product:
\begin{equation}
  f^U_{ab}(g)f^V_{cd}(g)
  =
  D_{U\otimes V}(g)_{(a,c),(b,d)} .
  \label{eq:dr-s3-matrix-coefficient-product}
\end{equation}
Decomposing \(U\otimes V\) by intertwiners gives the fusion rules such as
\eqref{eq:dr-s3-representation-ring}.  This finite model is not a local QFT
field algebra; its observable part is only \(\mathbb C\).  Its purpose is to
show, without the analytic burden of DHR localization, how nonabelian charged
coordinates, Haar expectation, and fusion multiplication are the same data
that the Doplicher--Roberts theorem reconstructs from a symmetric
finite-statistics sector category.

\paragraph{Finite Tannakian converse through the regular algebra.}
The finite regular algebra also makes the converse direction of the
Tannakian reconstruction visible.  Let \(G\) be a finite group and let
\[
  \mathcal R=\operatorname{Fun}(G)
\]
with point idempotents \(\delta_x\), \(x\in G\).  The left regular action is
\begin{equation}
  (L_a f)(g)=f(a^{-1}g),
  \qquad
  L_a\delta_x=\delta_{ax}.
  \label{eq:finite-tannaka-left-regular-action}
\end{equation}
Every unital \(*\)-automorphism \(T\) of the commutative finite-dimensional
algebra \(\mathcal R\) permutes the minimal projections:
\[
  T\delta_x=\delta_{\phi(x)}
\]
for a unique permutation \(\phi\) of \(G\).  If \(T\) commutes with every
left translation, then
\[
  \delta_{\phi(ax)}
  =
  T L_a\delta_x
  =
  L_a T\delta_x
  =
  \delta_{a\phi(x)} .
\]
Hence \(\phi(ax)=a\phi(x)\).  Setting \(x=e\) gives
\[
  \phi(a)=a h,\qquad h=\phi(e),
\]
and therefore \(\phi(x)=xh\) for all \(x\in G\).  Thus every
left-equivariant algebra automorphism of \(\operatorname{Fun}(G)\) is a right
translation:
\begin{equation}
  (T_h f)(g)=f(g h^{-1}).
  \label{eq:finite-tannaka-right-translation-classification}
\end{equation}

For a matrix coefficient \(f^W_{\alpha i}(g)=D_W(g)_{\alpha i}\),
\eqref{eq:finite-tannaka-right-translation-classification} gives
\begin{equation}
  T_h f^W_{\alpha i}(g)
  =
  D_W(g h^{-1})_{\alpha i}
  =
  \sum_j f^W_{\alpha j}(g)\,D_W(h^{-1})_{ji}.
  \label{eq:finite-tannaka-matrix-coefficient-action}
\end{equation}
Thus the corresponding tensor natural automorphism of the forgetful fiber
functor is
\[
  u_W=D_W(h^{-1})
\]
on every finite-dimensional representation \(W\).  Conversely, a unitary
tensor natural automorphism \(u_W\) acts on matrix coefficients by the right
side of \eqref{eq:finite-tannaka-matrix-coefficient-action}; tensor
compatibility makes this action multiplicative under
\eqref{eq:dr-s3-matrix-coefficient-product}, naturality makes it commute
with the intertwiners that decompose tensor products, and the Peter--Weyl
decomposition of \(\operatorname{Fun}(G)\) gives a left-equivariant
\(*\)-automorphism of \(\mathcal R\).  The preceding paragraph then supplies a
unique \(h\in G\).  For finite groups this proves
\[
  \operatorname{Aut}^{\otimes,*}(\omega_{\rm fib})
  \simeq
  G
\]
in the convention \(u_W=D_W(h^{-1})\).  The compact-group
Doplicher--Roberts theorem uses the same Peter--Weyl mechanism with the
dense Hopf \(*\)-algebra of representative functions and the \(C^*\)-norm
closure; the local-QFT burden remains the DHR construction of the symmetric
finite-statistics sector category and the field-algebra realization of this
categorical duality.

\paragraph{Compact Peter--Weyl coordinate algebra.}
\label{par:compact-peter-weyl-coordinate-algebra}
The compact version of the preceding finite calculation is the
representative-function Hopf \(*\)-algebra.  Let \(G\) be a compact group and
let \(W\) range over finite-dimensional continuous unitary representations.
For \(v\in W\) and \(\ell\in W^*\), write
\[
  f_{\ell,v}^W(g)=\ell(D_W(g)v).
\]
The algebra
\[
  \operatorname{Pol}(G)
  =
  \operatorname{span}\{f_{\ell,v}^W\}
  \subset C(G)
\]
is closed under multiplication because
\begin{equation}
  f_{\ell_1,v_1}^{W_1}(g)\,f_{\ell_2,v_2}^{W_2}(g)
  =
  f_{\ell_1\otimes\ell_2,\,v_1\otimes v_2}^{W_1\otimes W_2}(g),
  \label{eq:peter-weyl-product-tensor}
\end{equation}
and it is closed under \(*\) because complex conjugation is the matrix
coefficient of the conjugate representation.  The Hopf operations are the
group-law operations written on functions; in this paragraph write
\(\Delta_{\rm H}\) for the coproduct, to avoid confusing it with the modular
operator \(\Delta\) used elsewhere in the chapter:
\begin{equation}
\begin{aligned}
  (\Delta_{\rm H} f)(g,h)&=f(gh),
&
  \epsilon(f)&=f(e),
&
  (Sf)(g)&=f(g^{-1}).
\end{aligned}
  \label{eq:peter-weyl-hopf-operations}
\end{equation}
Associativity of group multiplication gives coassociativity of
\(\Delta_{\rm H}\); the identity element gives the counit identities;
inversion gives the antipode identities
\[
  m(S\otimes\operatorname{id})\Delta_{\rm H} f
  =
  m(\operatorname{id}\otimes S)\Delta_{\rm H} f
  =
  \epsilon(f)\mathbf1 .
\]
Peter--Weyl says that \(\operatorname{Pol}(G)\) is uniformly dense in \(C(G)\),
and Schur orthogonality decomposes it as the algebraic direct sum of matrix
coefficient spaces
\[
  \operatorname{Pol}(G)
  \simeq
  \bigoplus_{\widehat G} W^*\otimes W .
\]
Haar integration is the unique bi-invariant positive normalized functional
\[
  h_G(f)=\int_G f(g)\,\dd g ,
\]
and it kills every nontrivial irreducible matrix-coefficient block while
keeping the trivial block.

This compact Peter--Weyl algebra is the coordinate side of the compact group
appearing in Doplicher--Roberts reconstruction.  The tensor category supplies
the representations \(W\), tensor products supply multiplication in
\eqref{eq:peter-weyl-product-tensor}, and tensor natural automorphisms of the
fiber functor are evaluated by the same matrices \(D_W(g)\).  The
representative functions separate points of \(G\), so the compact group is
recoverable from how all charged multiplets transform; equivalently, after
the \(C^*\)-completion \(C(G)\) is supplied, the character space recovers
\(G\) with its compact topology.  None of this proves that a given observable
net has the required DHR category; it identifies the compact-group algebra
that appears after those local-QFT hypotheses have already been established.

\paragraph{Compact abelian completion: the \(U(1)\) charge lattice.}
The finite pointed example also has a compact limit that contains no
finite-group artifact.  Let \({\rm Rep}_{\rm fd}(U(1))\) denote the category
of finite-dimensional continuous unitary representations of \(U(1)\).  Its
simple objects are the one-dimensional characters \(L_n\), \(n\in\mathbb Z\),
with
\[
  L_m\otimes L_n\simeq L_{m+n},
  \qquad
  L_0=\mathbf 1,
  \qquad
  \operatorname{Hom}(L_m,L_n)=
  \begin{cases}
    \mathbb C, & m=n,\\
    0, & m\ne n .
  \end{cases}
\]
For the forgetful fiber functor
\(\omega_{\rm fib}:{\rm Rep}_{\rm fd}(U(1))\to{\rm Hilb}_{\rm fd}\), a unitary
tensor natural automorphism is a family of phases
\[
  u_n\in U(1),
  \qquad n\in\mathbb Z,
\]
satisfying
\[
  u_0=1,
  \qquad
  u_{m+n}=u_m u_n,
  \qquad
  u_{-n}=\overline{u_n}.
\]
The first two equations give \(u_n=u_1^n\) for \(n>0\) and then, by unitarity,
for all \(n\in\mathbb Z\).  Conversely every \(\lambda\in U(1)\) gives such a
family by \(u_n=\lambda^n\).  With the topology of pointwise convergence on
objects, the evaluation map
\[
  U(1)\longrightarrow \operatorname{Aut}^{\otimes,*}(\omega_{\rm fib}),
  \qquad
  \lambda\longmapsto (u_n=\lambda^n)_{n\in\mathbb Z},
\]
is therefore a homeomorphism: continuity is pointwise continuity of
\(\lambda^n\), and the inverse reads off \(u_1\).

The corresponding representative-function algebra is the Laurent polynomial
\(*\)-algebra
\[
  \mathbb C[z,z^{-1}],
  \qquad
  z^m z^n=z^{m+n},
  \qquad
  (z^n)^*=z^{-n}.
\]
Its \(C^*\)-completion in the supremum norm on the unit circle is
\[
  C(U(1)),
\]
because Laurent polynomials are a unital self-adjoint subalgebra separating
points of \(U(1)\).  Haar expectation is the projection onto the trivial
isotypic component:
\begin{equation}
  E_{U(1)}(z^n)
  =
  \int_{U(1)} z^n\,\dd z
  =
  \delta_{n0}.
  \label{eq:u1-tannaka-haar-laurent}
\end{equation}
Thus a \(U(1)\)-graded charged algebraic core
\[
  \mathcal F_{\rm alg}
  =
  \bigoplus_{n\in\mathbb Z}\mathcal F_n,
  \qquad
  \mathcal F_m\mathcal F_n\subset\mathcal F_{m+n},
  \qquad
  \alpha_\lambda(F_n)=\lambda^nF_n
\]
has invariant part
\[
  \mathcal F_{\rm alg}^{U(1)}=\mathcal F_0
\]
on every finite Laurent-polynomial subspace.  This compact abelian
calculation is the charge-lattice version of the finite
\(\mathbb Z/N\mathbb Z\) diagnostic above.  It supplies the categorical and
representative-function completion, not the local-QFT theorem: a concrete
QFT still has to produce the localized symmetric finite-statistics category,
the field algebra, and the action whose fixed points are the observable net.

\begin{example}[Pointed field algebra as a crossed product]
\label{ex:dr-pointed-crossed-product-field-algebra}
The same pointed category has an explicit charged-field algebra when the
sector \(X_1\) is represented by an order-\(N\) localized automorphism
\[
  \rho\in\operatorname{Aut}(\mathcal A),
  \qquad
  \rho^N=\operatorname{id}_{\mathcal A},
\]
of a neutral algebra \(\mathcal A\).  Introduce a unitary charged generator
\(u\) with
\[
  u^N=1,\qquad uAu^*=\rho(A),\quad A\in\mathcal A,
\]
and define the algebraic field core
\begin{equation}
  \mathcal F_{\rm alg}
  =
  \bigoplus_{q\in\mathbb Z/N\mathbb Z}\mathcal A u^q .
  \label{eq:dr-pointed-crossed-product-core}
\end{equation}
The multiplication and involution are
\begin{equation}
  (Au^q)(Bu^r)
  =
  A\,\rho^q(B)\,u^{q+r},
  \qquad
  (Au^q)^*
  =
  \rho^{-q}(A^*)u^{-q}.
  \label{eq:dr-pointed-crossed-product-operations}
\end{equation}
Associativity follows from the group law \(\rho^{q+r}=\rho^q\rho^r\):
\[
\begin{aligned}
  ((Au^q)(Bu^r))(Cu^s)
  &=
  A\rho^q(B)\rho^{q+r}(C)u^{q+r+s},\\
  (Au^q)((Bu^r)(Cu^s))
  &=
  A\rho^q\!\bigl(B\rho^r(C)\bigr)u^{q+r+s}.
\end{aligned}
\]
The two expressions agree because \(\rho^q\) is an algebra automorphism.  The
involution is anti-multiplicative because
\[
\begin{aligned}
  \bigl((Au^q)(Bu^r)\bigr)^*
  &=
  \rho^{-q-r}\!\bigl(\rho^q(B^*)A^*\bigr)u^{-q-r}  \\
  &=
  \rho^{-r}(B^*)\,\rho^{-q-r}(A^*)u^{-q-r}
  =
  (Bu^r)^*(Au^q)^* .
\end{aligned}
\]
The reconstructed compact group
\(\mu_N=\{\zeta\in U(1):\zeta^N=1\}\) acts by
\begin{equation}
  \alpha_\zeta(Au^q)=\zeta^q Au^q .
  \label{eq:dr-pointed-crossed-product-gauge-action}
\end{equation}
Its fixed algebra is exactly the neutral algebra:
\[
  \mathcal F_{\rm alg}^{\mu_N}=\mathcal A .
\]
Indeed, the Haar average over the finite group is
\[
  E_{\mu_N}\!\left(\sum_q A_q u^q\right)
  =
  \frac1N\sum_{\zeta^N=1}\sum_q \zeta^q A_q u^q
  =
  A_0 .
\]
Thus the charge creator \(u\) implements the localized sector by
\[
  \rho(A)=uAu^*,
\]
while the observable algebra is recovered as the invariant subalgebra.  This
example is the DR field-algebra output in the invertible pointed case.  It is
an algebraic output construction.  To interpret it as a local charged-field
net, the QFT must first supply the DHR input package: locality,
transportability, Haag duality, and finite statistics for the sector
endomorphisms.
\end{example}

This theorem therefore applies to charges whose sector representatives are
localized in bounded regions and satisfy the finite-statistics DHR
hypotheses.  Charged sectors controlled by long-range Gauss laws, by
Wilson-line or spacelike-cone localization, by infraparticle clouds, or by
confining dynamics require different localization categories and additional
analytic input.  Such sectors are not failures of the reconstruction theorem;
they lie outside the DHR category to which the theorem is attached.  For them
one must first define the relevant charged operators and localization
structure before asking for an analogue of field-algebra reconstruction or of
Haag--Ruelle scattering.

\paragraph{Constructive kink-sector diagnostic.}
The boundary of the DHR category is visible in broken two-dimensional
constructive scalar models.  Suppose a massive model has pure phases indexed
by \(\sigma\), vacuum representations \(\pi_\sigma\), and a bounded local
order-parameter observable \(B\) with distinct phase values \(m_\sigma\).
More precisely, choose spatial averaging functions supported in the far left
and far right tails and let \(B_L(R)\), \(B_R(R)\) denote the corresponding
bounded observables, with supports moving to \(x^1<-R\) and \(x^1>R\).  A
solitonic representation
\[
  \pi_{\sigma_-,\sigma_+}
\]
interpolating from phase \(\sigma_-\) at \(x^1=-\infty\) to phase
\(\sigma_+\) at \(x^1=+\infty\) is characterized, when the constructive
limits have been proved, by weak-operator tail limits
\begin{equation}
  \mathop{\mathrm{w\!-\!lim}}\limits_{R\to\infty}
  \pi_{\sigma_-,\sigma_+}(B_L(R))
  =
  m_{\sigma_-}\mathbf 1,\qquad
  \mathop{\mathrm{w\!-\!lim}}\limits_{R\to\infty}
  \pi_{\sigma_-,\sigma_+}(B_R(R))
  =
  m_{\sigma_+}\mathbf 1 .
  \label{eq:kink-tail-order-parameter-limits}
\end{equation}
These limits are macroscopic phase data, not local charged-field data.  If
\(\pi_{\sigma_-,\sigma_+}\) were DHR-localized in a bounded double cone
\(\mathcal O\) relative to a vacuum phase \(\pi_{\sigma_0}\), then for
sufficiently large \(R\) both \(B_L(R)\) and \(B_R(R)\) would belong to
\(\Obs(\mathcal O^\perp)\).  The DHR equivalence
\[
  \pi_{\sigma_-,\sigma_+}|_{\Obs(\mathcal O^\perp)}
  \simeq
  \pi_{\sigma_0}|_{\Obs(\mathcal O^\perp)}
\]
would therefore carry the weak limits of the two tails to the corresponding
tail limits in the phase \(\sigma_0\), namely \(m_{\sigma_0}\mathbf 1\) on
both sides.  Hence bounded-region DHR localization forces
\[
  m_{\sigma_-}=m_{\sigma_0}=m_{\sigma_+}.
\]
Whenever the left and right asymptotic phases differ, the soliton sector is
therefore not a DHR sector relative to a single vacuum representation.  It is
a sector controlled by asymptotic boundary conditions, equivalently by
morphisms between phase representations or by half-line/spacelike-cone
localization data.  The diagnostic does not assert the existence of such
soliton Hilbert spaces; that is a model theorem.  It shows which
representation-theoretic statement must be proved before a constructive kink
can be brought into the DHR local-endomorphism framework, and why the
Doplicher--Roberts theorem above is a bounded-charge theorem rather than a
classification of all superselection sectors in low-dimensional models.

\section{Split Property and Local Tensor Factorization}

Let \(\mathcal O_1\Subset \mathcal O_2\) mean that the closure of
\(\mathcal O_1\) is contained in \(\mathcal O_2\) with a positive spacetime
collar.  The split property is the existence
of a type I factor \(\mathcal N\) satisfying
\begin{equation}
  \mathcal R(\mathcal O_1)\subset \mathcal N\subset \mathcal R(\mathcal O_2).
  \label{eq:split-inclusion}
\end{equation}
A type I factor is isomorphic to \(\mathcal B(\mathcal K)\) for some Hilbert space
\(\mathcal K\).  The split inclusion gives a controlled tensor-product
separation between observables well inside \(\mathcal O_1\) and observables
outside \(\mathcal O_2\).  It is a precise algebraic substitute for a
finite-dimensional
subsystem factorization at nonzero separation.

Figure~\ref{fig:split-property} records the split inclusion
\(\mathcal R(\mathcal O_1)\subset\mathcal N\subset\mathcal R(\mathcal O_2)\)
and its role as a type-I tensor-product separation at positive distance.

\begin{figure}[H]
  \centering
  \resizebox{0.86\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      alg/.style={draw,rounded corners=4pt,line width=0.8pt,fill=blue!4,
        minimum width=2.9cm,minimum height=0.85cm,align=center},
      arrow/.style={-{Latex[length=2mm]},line width=0.85pt}
    }
    \draw[rounded corners=8pt,line width=0.9pt,fill=orange!5]
      (-4.8,-1.15) rectangle (4.8,1.15);
    \draw[rounded corners=8pt,line width=0.9pt,fill=white]
      (-2.1,-0.65) rectangle (2.1,0.65);
    \node at (3.75,0.72) {\(\mathcal O_2\)};
    \node at (1.5,0.28) {\(\mathcal O_1\)};
    \node[alg] (r1) at (-3.45,-1.9) {\(\mathcal R(\mathcal O_1)\)};
    \node[alg,fill=green!7] (n) at (0,-1.9) {type I factor\\\(\mathcal N\)};
    \node[alg] (r2) at (3.45,-1.9) {\(\mathcal R(\mathcal O_2)\)};
    \draw[arrow] (r1)--(n);
    \draw[arrow] (n)--(r2);
    \node at (0,-2.75)
      {\(\mathcal R(\mathcal O_1)\subset\mathcal N\subset\mathcal R(\mathcal O_2)\)};
  \end{tikzpicture}
  }
  \caption{The split property inserts a type I factor between nested local
  algebras separated by a collar.}
  \label{fig:split-property}
\end{figure}

The split property is tied to phase-space and nuclearity conditions.  Such
conditions control the number of local degrees of freedom below a given energy
scale and make local thermodynamic behavior mathematically tractable.

The operational meaning of the type-I interpolant is normal statistical
independence at positive separation.  If
\(\mathcal M_1=\mathcal R(\mathcal O_1)\),
\(\mathcal M_2=\mathcal R(\mathcal O_2)\), and
\(\mathcal M_1\subset\mathcal N\subset\mathcal M_2\) is split, choose a
spatial representation
\[
  U\mathcal N U^*=\mathcal B(\mathcal K)\otimes\mathbf 1,
  \qquad
  U\mathcal N'U^*=\mathbf 1\otimes\mathcal B(\mathcal L).
\]
Then \(U\mathcal M_1U^*\subset\mathcal B(\mathcal K)\otimes\mathbf 1\) and
\(U\mathcal M_2'U^*\subset\mathbf 1\otimes\mathcal B(\mathcal L)\).  Hence
normal states \(\varphi_1\) on \(\mathcal M_1\) and \(\varphi_2\) on
\(\mathcal M_2'\) extend to normal states
\(\widetilde\varphi_1\) on \(\mathcal B(\mathcal K)\) and
\(\widetilde\varphi_2\) on \(\mathcal B(\mathcal L)\).  The normal product
state
\(\widetilde\varphi_1\bar\otimes\widetilde\varphi_2\) restricts to a normal
state on the von Neumann algebra generated by the two commuting algebras, and
on products it satisfies
\[
  \Phi(AB')
  =
  \varphi_1(A)\varphi_2(B'),
  \qquad
  A\in\mathcal M_1,\quad B'\in\mathcal M_2' .
\]
The construction uses the ordinary tensor product after conjugating by \(U\);
the positive collar is what permits this tensor product to be represented
normally inside the vacuum Hilbert space.  As the collar is removed, the
interpolating type-I factor is no longer part of the sharp local-algebra data,
which is why split limits and relative entropy are the intrinsic continuum
objects used later.

\section{Modular Data and Locality}

Let \(\mathcal R\) be a von Neumann algebra on a Hilbert space with a vector
\(\vac\) that is cyclic and separating for \(\mathcal R\).  Tomita--Takesaki
theory defines an antilinear operator \(S\) by
\[
  S A\vac=A^*\vac,\qquad A\in\mathcal R,
\]
and its polar decomposition
\[
  S=J\Delta^{1/2}.
\]
Here \(J\) is antiunitary and \(\Delta\) is positive self-adjoint.  The
operator \(\Delta\) and its square root may be unbounded, but the Borel
functional calculus for positive self-adjoint operators gives
\(\Delta^{\ii t}\) as a bounded unitary operator for every \(t\in\mathbb R\),
forming a strongly continuous one-parameter unitary group.  The modular
automorphism group is
\[
  \sigma_t(A)=\Delta^{\ii t}A\Delta^{-\ii t}.
\]

\begin{quotedtheorem}[Tomita--Takesaki modular automorphisms]
\emph{Proof references.}\footnote{A complete proof of Tomita's theorem and
the modular automorphism theorem is given in M.~Takesaki,
\emph{Tomita's Theory of Modular Hilbert Algebras and Its Applications},
Lecture Notes in Mathematics \textbf{128}, Springer, 1970.  See also
M.~Takesaki, \emph{Theory of Operator Algebras I}, Springer, 1979, Chapter
VI, for the von Neumann algebra formulation used here \cite{Takesaki1970}.}
For a cyclic and separating vector \(\vac\) for \(\mathcal R\), the modular
operators constructed above satisfy, for every \(t\in\mathbb R\),
\[
  \Delta^{\ii t}\mathcal R\Delta^{-\ii t}=\mathcal R .
\]
Consequently
\[
  \sigma_t=\operatorname{Ad}_{\Delta^{\ii t}}\big|_{\mathcal R}
\]
is a one-parameter group of \(*\)-automorphisms of \(\mathcal R\).  Moreover,
the modular conjugation identifies the algebra with its commutant:
\[
  J\mathcal R J=\mathcal R',
\]
where \(\mathcal R'\) denotes the commutant of \(\mathcal R\) in
\(\mathcal B(\Hilb)\).
\end{quotedtheorem}

\begin{example}[Finite standard form of a faithful matrix state]
\label{ex:finite-tomita-standard-form}
Let \(K\simeq\mathbb C^n\), let
\(\mathcal R=\mathcal B(K)\), and let
\(\rho\in\mathcal B(K)\) be a positive definite matrix with
\({\rm Tr}_K\rho=1\).  Put
\[
  \Hilb_{\rm HS}=\mathcal B(K),
  \qquad
  \langle X,Y\rangle_{\rm HS}={\rm Tr}_K(X^*Y),
  \qquad
  \pi(A)X=AX,
  \qquad
  \Omega=\rho^{1/2}.
\]
Then \(\Omega\) is cyclic and separating for \(\pi(\mathcal R)\).  The
Tomita operator, modular conjugation, and modular operator are
\begin{equation}
  S(X)=\rho^{-1/2}X^*\rho^{1/2},
  \qquad
  J(X)=X^*,
  \qquad
  \Delta(X)=\rho X\rho^{-1}.
  \label{eq:finite-standard-form-tomita-data}
\end{equation}
Consequently
\[
  J\pi(\mathcal R)J=\pi(\mathcal R)',
  \qquad
  \sigma_t(A)=\rho^{\ii t}A\rho^{-\ii t}.
\]
With the boundary-value convention used below for KMS states,
\(F(t)=\omega(A\alpha_t(B))\) and
\(F(t+\ii)=\omega(\alpha_t(B)A)\), the vector state
\[
  \omega(A)=\langle\Omega,\pi(A)\Omega\rangle_{\rm HS}
  ={\rm Tr}_K(\rho A)
\]
is a \(1\)-KMS state for \(\alpha_t=\sigma_{-t}\).
\end{example}

\paragraph{Verification of the finite model.}
Every \(X\in\Hilb_{\rm HS}\) has the form
\[
  X=(X\rho^{-1/2})\Omega,
\]
so \(\Omega\) is cyclic.  If \(A\Omega=0\), then
\(A\rho^{1/2}=0\), and invertibility of \(\rho^{1/2}\) gives \(A=0\);
hence \(\Omega\) is separating.  For \(X=A\Omega\),
\[
  S(X)=S(A\rho^{1/2})=A^*\rho^{1/2}
  =
  \rho^{-1/2}X^*\rho^{1/2}.
\]
The Hilbert-Schmidt basis of matrix units diagonalizing \(\rho\) shows that
\[
  (J\Delta^{1/2})(X)
  =
  J(\rho^{1/2}X\rho^{-1/2})
  =
  \rho^{-1/2}X^*\rho^{1/2}=S(X),
\]
with \(J(X)=X^*\) and \(\Delta=L_\rho R_{\rho^{-1}}\), where \(L\) and
\(R\) denote left and right multiplication.  The commutant of all left
multiplications is the algebra of right multiplications, and
\[
  J\pi(A)J(X)=J(A X^*)=X A^*,
\]
so \(J\pi(\mathcal R)J=\pi(\mathcal R)'\).  Conjugating a left
multiplication by \(\Delta^{\ii t}\) gives
\[
  \Delta^{\ii t}\pi(A)\Delta^{-\ii t}X
  =
  \rho^{\ii t}A\rho^{-\ii t}X,
\]
which is the displayed modular automorphism.

For this boundary-value convention, define
\[
  F_{A,B}(z)
  =
  {\rm Tr}_K\!\left(\rho A\rho^{-\ii z}B\rho^{\ii z}\right),
  \qquad 0\leq{\rm Im}\,z\leq1 .
\]
This is an entire finite sum after diagonalizing \(\rho\).  The lower
boundary is \(F_{A,B}(t)=\omega(A\alpha_t(B))\).  At the upper boundary,
using \(\alpha_t(B)=\rho^{-\ii t}B\rho^{\ii t}\) and cyclicity of the finite
trace,
\[
\begin{aligned}
  F_{A,B}(t+\ii)
  &=
  {\rm Tr}_K\!\left(\rho A\rho^{-\ii t+1}B\rho^{\ii t-1}\right)  \\
  &=
  {\rm Tr}_K\!\left(A\rho\rho^{-\ii t}B\rho^{\ii t}\right)
  =
  {\rm Tr}_K\!\left(\rho\alpha_t(B)A\right)
  =
  \omega(\alpha_t(B)A).
\end{aligned}
\]
Thus \(\omega\) satisfies the \(1\)-KMS boundary condition for
\(\alpha_t=\sigma_{-t}\).

\paragraph{Proof mechanism and local role.}
The theorem is a statement about a standard von Neumann algebra
\((\mathcal R,\Hilb,\vac)\), where standard means that \(\vac\) is cyclic and
separating for \(\mathcal R\).  The proof begins with the second antilinear
operator
\[
  F_0(B\vac)=B^*\vac,\qquad B\in\mathcal R' .
\]
The commutation of \(\mathcal R\) and \(\mathcal R'\) gives the adjoint
identity for antilinear operators on the two algebraic cores,
\[
  \langle S_0(A\vac),B\vac\rangle
  =
  \langle A^*\vac,B\vac\rangle
  =
  \langle B^*\vac,A\vac\rangle
  =
  \langle F_0(B\vac),A\vac\rangle ,
  \qquad A\in\mathcal R,\ B\in\mathcal R' ,
\]
using \(AB=BA\).  Hence \(S_0\subset F_0^*\) and \(F_0\subset S_0^*\).
Cyclicity of \(\vac\) for \(\mathcal R\) and \(\mathcal R'\) supplies dense
domains, and the two inclusions imply closability.  Denote the closures by
\(S\) and \(F\).  The same adjoint identity upgrades to
\[
  S^*=F,\qquad F^*=S .
\]
The polar decomposition of the closed antilinear operator \(S\) gives
\[
  S=J\Delta^{1/2},
  \qquad
  F=J\Delta^{-1/2},
  \qquad
  J\Delta J=\Delta^{-1}.
\]

The substantive step is the Tomita invariance theorem for the left Hilbert
algebra \(\mathfrak A=\mathcal R\vac\).  The product and involution on the
core are
\[
  (A\vac)(B\vac)=AB\vac,\qquad
  (A\vac)^\#=A^*\vac .
\]
Left multiplication by \(A\in\mathcal R\) is bounded on the Hilbert-space
completion.  The closure of the involution is precisely \(S\).  Tomita's
analytic argument shows that the closed left Hilbert algebra is stable under
the modular unitaries \(\Delta^{\ii t}\), and therefore
\(\Delta^{\ii t}A\Delta^{-\ii t}\) is again a bounded left multiplier of
\(\mathfrak A\).  Bounded left multipliers of the closed algebra are exactly
the elements of \(\mathcal R\).  This proves
\[
  \Delta^{\ii t}\mathcal R\Delta^{-\ii t}\subset \mathcal R .
\]
Applying the same argument at time \(-t\) gives equality.  The commutant
identity follows from the polar part of the same closed involution: \(J A J\)
acts as right multiplication by \(A^*\) on the core and hence commutes with
all left multipliers, while the symmetric argument starting from
\(\mathcal R'\) gives the reverse inclusion.

For the QFT applications in this chapter the theorem supplies an abstract
modular flow for any local algebra whose vacuum vector is standard.  The QFT
work begins after this theorem: one must prove Reeh--Schlieder or another
standardness criterion for the local algebra, and one must separately identify
the abstract modular flow with a geometric flow when such an identification is
claimed.  The Bisognano--Wichmann theorem boundary below is precisely that
additional QFT input for wedges.

\section{KMS States and Modular Equilibrium}
\label{sec:kms-states-modular-equilibrium}

The modular group is not only an algebraic automorphism group.  It is the
canonical source of the operator-algebraic equilibrium condition.  We first
state the condition for an arbitrary time evolution, because this is the form
used in finite-temperature QFT and in infinite-volume statistical mechanics.

\begin{definition}[KMS state]
\label{def:kms-state}
Let \(\mathfrak A\) be a \(C^*\)-algebra or a von Neumann algebra, let
\(\alpha:\mathbb R\to\operatorname{Aut}(\mathfrak A)\) be a strongly
continuous one-parameter automorphism group, and let \(\omega\) be a state on
\(\mathfrak A\).  For \(\beta>0\), \(\omega\) is a
\(\beta\)-KMS state for \(\alpha\) when, for every pair
\(A,B\in\mathfrak A\), there is a function
\[
  F_{A,B}:\{z\in\mathbb C:0\le \operatorname{Im}z\le\beta\}\to\mathbb C
\]
which is bounded and continuous on the closed strip, holomorphic in its
interior, and satisfies the boundary conditions
\begin{equation}
  F_{A,B}(t)=\omega\!\left(A\,\alpha_t(B)\right),
  \qquad
  F_{A,B}(t+\ii\beta)=
  \omega\!\left(\alpha_t(B)\,A\right),
  \qquad t\in\mathbb R .
  \label{eq:kms-boundary-condition}
\end{equation}
\end{definition}

The order of the two boundary values is the content of the condition.  It is
not periodicity of an ordinary function of real time; it is an analytic
continuation statement whose upper boundary has the operators cyclically
permuted.  This distinction is essential in infinite volume, where the trace
of \(\ee^{-\beta H}\) usually does not exist but the local state may still
satisfy \eqref{eq:kms-boundary-condition}.

\begin{example}[Gibbs traces satisfy KMS]
\label{ex:gibbs-trace-kms}
Let \(H\) be self-adjoint and suppose that
\(\ee^{-\beta H}\) is trace class.  On the algebra of bounded operators for
which the following expressions are defined by trace-class products, set
\[
  \omega_\beta(A)
  =
  Z_\beta^{-1}\operatorname{Tr}\left(\ee^{-\beta H}A\right),
  \qquad
  Z_\beta=\operatorname{Tr}\ee^{-\beta H},
  \qquad
  \alpha_t(A)=\ee^{\ii tH}A\ee^{-\ii tH}.
\]
Then \(\omega_\beta\) is a \(\beta\)-KMS state for \(\alpha\).
\end{example}

\paragraph{Verification of the trace-class KMS strip.}
For bounded \(A,B\), define
\[
  F_{A,B}(z)
  =
  Z_\beta^{-1}
  \operatorname{Tr}\left(
    \ee^{-\beta H}A\ee^{\ii zH}B\ee^{-\ii zH}
  \right).
\]
In finite dimension this is entire.  Under the stated trace-class hypothesis
the same formula is holomorphic in the open strip and bounded on the closed
strip by the standard trace-norm estimates for products with
\(\ee^{-sH}\), \(0\le s\le\beta\); finite-dimensional spectral truncations
give the same identity and converge in trace norm.  At the lower boundary,
\[
  F_{A,B}(t)=\omega_\beta(A\alpha_t(B)).
\]
At the upper boundary,
\[
\begin{aligned}
  F_{A,B}(t+\ii\beta)
  &=
  Z_\beta^{-1}
  \operatorname{Tr}\left(
    \ee^{-\beta H}A\ee^{\ii tH}\ee^{-\beta H}
    B\ee^{-\ii tH}\ee^{\beta H}
  \right)  \\
  &=
  Z_\beta^{-1}
  \operatorname{Tr}\left(
    A\ee^{\ii tH}\ee^{-\beta H}B\ee^{-\ii tH}
  \right)  \\
  &=
  Z_\beta^{-1}
  \operatorname{Tr}\left(
    \ee^{-\beta H}\ee^{\ii tH}B\ee^{-\ii tH}A
  \right)
  =
  \omega_\beta(\alpha_t(B)A).
\end{aligned}
\]
The first equality is the analytic continuation.  The second uses cyclicity of
the trace to move the final \(\ee^{\beta H}\) to the front, where it cancels
one \(\ee^{-\beta H}\).  The third uses cyclicity again and the commutation of
\(\ee^{-\beta H}\) with \(\ee^{\ii tH}\).  These two trace-cyclicity moves are
precisely what the KMS boundary condition retains after the literal Gibbs
trace ceases to exist.

\begin{theorem}[Modular KMS property]
\label{thm:modular-kms-property}
Let \(\mathcal R\) be a von Neumann algebra and let \(\vac\) be cyclic and
separating for \(\mathcal R\).  Let
\[
  \omega(A)=\langle\vac,A\vac\rangle,
  \qquad
  \sigma_t(A)=\Delta^{\ii t}A\Delta^{-\ii t}
\]
be the vector state and the Tomita--Takesaki modular automorphism group.
Then \(\omega\) satisfies the unit-width strip KMS condition for the modular
flow.  Equivalently, for every \(A,B\in\mathcal R\) there is a bounded strip
function \(G_{A,B}\), analytic for \(0<\operatorname{Im}z<1\), such that
\begin{equation}
  G_{A,B}(t)=\omega(\sigma_t(A)B),
  \qquad
  G_{A,B}(t+\ii)=\omega(B\sigma_t(A)).
  \label{eq:modular-kms-boundary}
\end{equation}
In the convention of Definition~\ref{def:kms-state}, this says that
\(\omega\) is a \(1\)-KMS state for the reversed modular dynamics
\(\alpha_t=\sigma_{-t}\).
\end{theorem}

\begin{proof}
The nontrivial input is Tomita--Takesaki theory: the closable antilinear
operator \(S_0A\vac=A^*\vac\) has polar decomposition
\(S=J\Delta^{1/2}\), the unitary group \(\Delta^{\ii t}\) implements
automorphisms of \(\mathcal R\), and the analytic vectors for this unitary
group form a strong-operator dense \(*\)-subalgebra of \(\mathcal R\).  For
analytic \(A\), the vector-valued function
\(z\mapsto\sigma_z(A)\vac\) is analytic in the strip.  Pairing it with
\(B^*\vac\) gives the candidate \(G_{A,B}(z)\).  The lower boundary is the
definition of \(\sigma_t\).  The upper boundary follows from
\[
  S\,\sigma_t(A)\vac=\sigma_t(A)^*\vac,
  \qquad
  S=J\Delta^{1/2},
\]
or, equivalently, from moving a vector through the closed graph of \(S\):
the operation that is trace cyclicity in
Example~\ref{ex:gibbs-trace-kms} is replaced by the Tomita relation
between \(\mathcal R\) and \(\mathcal R'\).  Thus
\[
  \langle\vac,\sigma_{t+\ii}(A)B\vac\rangle
  =
  \langle\vac,B\sigma_t(A)\vac\rangle
\]
for analytic \(A\).  The general case follows by Gaussian smearing in modular
time,
\[
  A_\epsilon
  =
  \frac1{\sqrt{\pi\epsilon}}
  \int_{\mathbb R}\ee^{-s^2/\epsilon}\sigma_s(A)\,\dd s ,
\]
which produces entire analytic elements converging strongly to \(A\) and
uniformly controls the strip functions by \(\|A\|\,\|B\|\).  Taking the limit
gives \eqref{eq:modular-kms-boundary}.
\end{proof}

The theorem is the precise sense in which a cyclic separating vector is an
equilibrium state for its own modular time.  The modular Hamiltonian
\[
  K_{\rm mod}:=-\log\Delta
\]
is generally not the physical Hamiltonian and need not be an integral of a
local density.  When a geometric theorem identifies modular flow with a
spacetime symmetry, the abstract KMS condition becomes a physical thermal
statement for the corresponding observer or subregion.

\section{Relative Modular Operators and Connes Cocycles}
\label{sec:relative-modular-connes-cocycles}

The modular data of a single pair \((\mathcal R,\Omega)\) compare an algebra
with its commutant.  The modular data of two states compare two normal
statistical preparations on the same algebra.  This relative form is the
operator-algebraic replacement for density-matrix quotients in local QFT,
where local algebras are typically type \(\mathrm{III}\) and no local density
matrix exists.

\begin{definition}[Relative Tomita operator]
\label{def:relative-tomita-operator}
Let \(\mathcal R\subset\mathcal B(\Hilb)\) be a von Neumann algebra, and let
\(\Psi,\Omega\in\Hilb\) be cyclic and separating for \(\mathcal R\).  Define
the antilinear operator \(S_{\Psi|\Omega}^{(0)}\) on the dense domain
\(\mathcal R\Omega\) by
\[
  S_{\Psi|\Omega}^{(0)}A\Omega=A^*\Psi,
  \qquad A\in\mathcal R .
\]
It is closable.  Its closure is denoted \(S_{\Psi|\Omega}\), and
\[
  \Delta_{\Psi|\Omega}:=S_{\Psi|\Omega}^*S_{\Psi|\Omega}
\]
is the relative modular operator.  The polar decomposition is
\[
  S_{\Psi|\Omega}=J_{\Psi|\Omega}\Delta_{\Psi|\Omega}^{1/2},
\]
where \(J_{\Psi|\Omega}\) is a partial antiunitary with initial and final
closures determined by the supports of the two normal vector states.  Under
the cyclic-and-separating hypothesis the supports are \(1\), so
\(J_{\Psi|\Omega}\) is antiunitary.
\end{definition}

The order in the subscript is part of the convention.  The operator
\(\Delta_{\Psi|\Omega}\) has the state represented by \(\Psi\) on the left
and the state represented by \(\Omega\) on the right.  In a finite type-I
algebra this statement becomes literal.  Let
\(\mathcal R=\mathcal B(\mathcal K)\) act by left multiplication on the
Hilbert--Schmidt space
\[
  \Hilb_{\rm st}=\{X:\mathcal K\to\mathcal K:
  \operatorname{Tr}(X^*X)<\infty\},
  \qquad
  \langle X,Y\rangle_{\rm st}=\operatorname{Tr}(X^*Y).
\]
A faithful normal state with density matrix \(\rho>0\) is represented by
\(\Omega_\rho=\rho^{1/2}\).  For faithful density matrices
\(\rho_\psi,\rho_\omega\),
\begin{equation}
  \Delta_{\psi|\omega}(X)
  =
  \rho_\psi X\rho_\omega^{-1}.
  \label{eq:finite-relative-modular-operator}
\end{equation}
Indeed, for \(X=A\rho_\omega^{1/2}\),
\[
  S_{\psi|\omega}X=A^*\rho_\psi^{1/2}.
\]
The adjoint relation in the Hilbert--Schmidt inner product gives
\[
  S_{\psi|\omega}^*S_{\psi|\omega}
  (A\rho_\omega^{1/2})
  =
  \rho_\psi A\rho_\omega^{-1/2}
  =
  \rho_\psi(A\rho_\omega^{1/2})\rho_\omega^{-1},
\]
which is \eqref{eq:finite-relative-modular-operator} on a dense set and hence
by closure.  This calculation fixes the convention used below.

\begin{definition}[Araki relative entropy]
\label{def:araki-relative-entropy}
Let \(\psi,\omega\) be faithful normal states represented by cyclic separating
vectors \(\Psi,\Omega\).  The Araki relative entropy of \(\psi\) with respect
to \(\omega\) is
\begin{equation}
  S_{\mathcal R}(\psi\|\omega)
  =
  -\langle\Psi,\log\Delta_{\Omega|\Psi}\Psi\rangle,
  \label{eq:araki-relative-entropy}
\end{equation}
with the value \(+\infty\) if the logarithmic quadratic form is not finite or
if the support of \(\psi\) is not dominated by the support of \(\omega\).
\end{definition}

In the finite type-I model, \eqref{eq:finite-relative-modular-operator}
turns \eqref{eq:araki-relative-entropy} into the ordinary Umegaki formula:
\[
\begin{aligned}
  S_{\mathcal B(\mathcal K)}(\psi\|\omega)
  &=
  -\operatorname{Tr}\!\left[
    \rho_\psi^{1/2}
    \bigl(\log\rho_\omega-\log\rho_\psi\bigr)
    \rho_\psi^{1/2}
  \right]  \\
  &=
  \operatorname{Tr}\rho_\psi(\log\rho_\psi-\log\rho_\omega).
\end{aligned}
\]
The algebraic definition is therefore not a new thermodynamic convention; it
is the continuation of the density-matrix expression to algebras for which no
density matrix exists.

\begin{quotedtheorem}[Connes cocycle derivative]
\label{thm:connes-cocycle-derivative}
\emph{Proof references.}\footnote{The cocycle derivative and Radon--Nikodym
theorem for faithful normal weights are proved in A.~Connes,
``Une classification des facteurs de type \(\mathrm{III}\),''
\emph{Annales scientifiques de l'Ecole Normale Superieure} \textbf{6}
(1973), 133--252, and in Takesaki, \emph{Theory of Operator Algebras II},
Chapter VIII \cite{Connes1973}.}
Let \(\omega\) and \(\psi\) be faithful normal states on a von Neumann algebra
\(\mathcal R\).  In a standard representation, define
\[
  u_t=[D\psi:D\omega]_t
  =
  \Delta_{\psi|\omega}^{\ii t}\Delta_\omega^{-\ii t},
  \qquad t\in\mathbb R ,
\]
where \(\Delta_\omega=\Delta_{\Omega|\Omega}\).  Then \(u_t\in\mathcal R\),
\(u_t\) is unitary, and
\begin{equation}
  u_{t+s}=u_t\,\sigma_t^\omega(u_s),
  \qquad t,s\in\mathbb R .
  \label{eq:connes-cocycle-relation}
\end{equation}
Moreover the modular flows of the two states are related by
\begin{equation}
  \sigma_t^\psi(A)
  =
  u_t\,\sigma_t^\omega(A)\,u_t^*,
  \qquad A\in\mathcal R .
  \label{eq:connes-cocycle-implements-flow}
\end{equation}
\end{quotedtheorem}

\paragraph{Proof mechanism and local role.}
The cocycle theorem is the Radon--Nikodym theorem for faithful normal states
in standard form.  The standard-form data are
\[
  (\mathcal R,\Hilb,J,\mathcal P),
\]
where \(J\) is the modular conjugation and \(\mathcal P\) is the natural
positive cone.  Each faithful normal state has a unique vector representative
in \(\mathcal P\); write these representatives as \(\Omega\) for \(\omega\)
and \(\Psi\) for \(\psi\).  The relative Tomita operator is initially defined
on the dense core \(\mathcal R\Omega\) by
\[
  S_{\Psi|\Omega}(A\Omega)=A^*\Psi,\qquad A\in\mathcal R .
\]
Its closure has polar data
\[
  S_{\Psi|\Omega}=J_{\Psi|\Omega}\Delta_{\psi|\omega}^{1/2}.
\]
The positive operator \(\Delta_{\psi|\omega}\) is the noncommutative
Radon--Nikodym derivative at the modular level.  The crucial assertion is that
the relative modular unitaries differ from the reference modular unitaries by
a unitary belonging to the algebra:
\[
  \Delta_{\psi|\omega}^{\ii t}\Delta_\omega^{-\ii t}\in\mathcal R .
\]
One proves this by testing the product against \(\mathcal R'\Omega\), using
the relative Tomita adjoint relation to show that it commutes with the
commutant, and then applying \((\mathcal R')'=\mathcal R\).  The same
relative-modular calculation gives
\[
  \Delta_{\psi|\omega}^{\ii(t+s)}\Delta_\omega^{-\ii(t+s)}
  =
  \bigl(\Delta_{\psi|\omega}^{\ii t}\Delta_\omega^{-\ii t}\bigr)
  \Delta_\omega^{\ii t}
  \bigl(\Delta_{\psi|\omega}^{\ii s}\Delta_\omega^{-\ii s}\bigr)
  \Delta_\omega^{-\ii t},
\]
which is exactly the cocycle relation
\eqref{eq:connes-cocycle-relation}.  Conjugating \(A\in\mathcal R\) by the
relative modular unitary and inserting the preceding factorization gives
\eqref{eq:connes-cocycle-implements-flow}.

The theorem's role in QFT is to compare modular Hamiltonians associated with
different states while staying inside the same local algebra.  Relative
entropy, modular energy, and the algebraic proof of monotonicity use this
cocycle as the replacement for a density-matrix quotient in type-III local
algebras.  The finite type-I calculation below is a normalization check; the
operator-algebraic theorem supplies the type-III object used by local QFT.

\begin{example}[Finite standard-form Connes cocycle]
\label{ex:finite-standard-form-connes-cocycle}
Let \(\rho_\omega,\rho_\psi\) be positive definite density matrices on
\(\mathcal K\), and represent
\(\mathcal R=\mathcal B(\mathcal K)\) on the Hilbert--Schmidt standard
space by left multiplication.  With the convention
\eqref{eq:finite-relative-modular-operator}, the operator
\[
  \Delta_{\psi|\omega}^{\ii t}\Delta_\omega^{-\ii t}
\]
is left multiplication by
\begin{equation}
  u_t=\rho_\psi^{\ii t}\rho_\omega^{-\ii t}\in\mathcal B(\mathcal K).
  \label{eq:finite-connes-cocycle-density-formula}
\end{equation}
The family \(u_t\) is unitary in \(\mathcal R\), satisfies the Connes
cocycle identity
\[
  u_{t+s}=u_t\,\sigma_t^\omega(u_s),
  \qquad
  \sigma_t^\omega(A)=\rho_\omega^{\ii t}A\rho_\omega^{-\ii t},
\]
and implements the change of modular flow,
\[
  \sigma_t^\psi(A)=u_t\sigma_t^\omega(A)u_t^*,
  \qquad
  \sigma_t^\psi(A)=\rho_\psi^{\ii t}A\rho_\psi^{-\ii t}.
\]
\end{example}

\paragraph{Verification of the finite cocycle.}
On the Hilbert--Schmidt space, the relative and reference modular unitaries
act by
\[
  \Delta_{\psi|\omega}^{\ii t}X
  =
  \rho_\psi^{\ii t}X\rho_\omega^{-\ii t},
  \qquad
  \Delta_\omega^{-\ii t}X
  =
  \rho_\omega^{-\ii t}X\rho_\omega^{\ii t}.
\]
Their product is therefore
\[
  \Delta_{\psi|\omega}^{\ii t}\Delta_\omega^{-\ii t}X
  =
  \rho_\psi^{\ii t}\rho_\omega^{-\ii t}X
  =
  u_tX,
\]
which is left multiplication by the matrix in
\eqref{eq:finite-connes-cocycle-density-formula}.  Since the factors
\(\rho_\psi^{\ii t}\) and \(\rho_\omega^{-\ii t}\) are unitary, \(u_t\) is a
unitary element of \(\mathcal R\).

The cocycle law keeps the density matrices in their noncommuting order:
\[
\begin{aligned}
  u_t\sigma_t^\omega(u_s)
  &=
  \rho_\psi^{\ii t}\rho_\omega^{-\ii t}
  \rho_\omega^{\ii t}
  (\rho_\psi^{\ii s}\rho_\omega^{-\ii s})
  \rho_\omega^{-\ii t}  \\
  &=
  \rho_\psi^{\ii(t+s)}\rho_\omega^{-\ii(t+s)}
  =
  u_{t+s}.
\end{aligned}
\]
Likewise,
\[
  u_t\sigma_t^\omega(A)u_t^*
  =
  \rho_\psi^{\ii t}\rho_\omega^{-\ii t}
  \rho_\omega^{\ii t}A\rho_\omega^{-\ii t}
  \rho_\omega^{\ii t}\rho_\psi^{-\ii t}
  =
  \rho_\psi^{\ii t}A\rho_\psi^{-\ii t}.
\]
This is exactly \(\sigma_t^\psi(A)\).  The finite calculation shows the
algebraic object that the general theorem constructs: an algebra unitary
whose cocycle law records the change of state and keeps the implementer
inside \(\mathcal R\).

For local QFT this quoted theorem is a calculational device.  Changing the state in a
bounded region is encoded not by dividing density matrices, but by a unitary
cocycle inside the local algebra.  In perturbative algebraic QFT this is the
operator-algebraic structure behind relative Cauchy evolution and local
source deformations; in constructive models it is the object on which
monotonicity, convexity, and energy inequalities are formulated.

\section{Type III Local Algebras and Phase-Space Bounds}
\label{sec:type-iii-local-algebras-phase-space-bounds}

The split property introduced above is an inclusion
\(\mathcal R(\mathcal O_1)\subset N\subset\mathcal R(\mathcal O_2)\) through a
type-I factor for separated regions, while sharply localized tensor
factorization would be a stronger and generally unavailable structure.  For a
proper bounded region \(\mathcal O\), the physically expected local algebra
\(\mathcal R(\mathcal O)\) is not a type-I factor.  Under the standard
structural hypotheses satisfied by known relativistic models, it is a
type \(\mathrm{III}_1\) factor: it has no nonzero finite projections and no
faithful normal trace.  The absence of a trace is why local thermal and
entanglement quantities must be formulated with modular theory rather than
with local density matrices.

\begin{definition}[Buchholz--Wichmann nuclearity map]
\label{def:buchholz-wichmann-nuclearity-map}
Let \(U(t)=\ee^{\ii tH}\) be the vacuum time-translation group with
\(H\ge0\), and let \(\mathcal R(\mathcal O)\) be a bounded-region algebra.
For \(\beta>0\), define
\[
  \Theta_{\beta,\mathcal O}:
  \mathcal R(\mathcal O)\to\Hilb,
  \qquad
  \Theta_{\beta,\mathcal O}(A)=\ee^{-\beta H}A\vac .
\]
The net satisfies the Buchholz--Wichmann nuclearity condition on
\(\mathcal O\) if \(\Theta_{\beta,\mathcal O}\) is a nuclear map for every
\(\beta>0\) and its nuclear norm obeys a local phase-space growth estimate,
for example
\[
  \|\Theta_{\beta,\mathcal O}\|_1
  \le
  \exp\!\left(C_{\mathcal O}\beta^{-s_{\mathcal O}}\right)
  \quad\text{as }\beta\downarrow0
\]
for some finite constants \(C_{\mathcal O},s_{\mathcal O}\).
\end{definition}

Here nuclear means that
\[
  \Theta_{\beta,\mathcal O}(A)=
  \sum_{n=1}^{\infty}\ell_n(A)\xi_n,
  \qquad
  \sum_n\|\ell_n\|\,\|\xi_n\|<\infty ,
\]
for bounded linear functionals \(\ell_n\) on \(\mathcal R(\mathcal O)\) and
vectors \(\xi_n\in\Hilb\).  The condition is an infinite-dimensional
replacement for saying that only finitely many independent localized states
are visible below energy scale \(1/\beta\).  It is not an ultraviolet cutoff:
it is a phase-space compactness estimate in the continuum theory.

\paragraph{Free Fock phase-space benchmark.}
The growth form in Definition~\ref{def:buchholz-wichmann-nuclearity-map} is
visible already in the free bosonic Fock counting problem.  Let the spatial
dimension be \(s=D-1\ge1\), put a scalar field of mass \(m>0\) in a spatial
torus of side length \(L\), and write the one-particle energies as
\[
  \epsilon_n
  =
  \sqrt{\left(\frac{2\pi}{L}\right)^2|n|_2^2+m^2},
  \qquad
  n\in\mathbb Z^s .
\]
On the bosonic Fock space the finite-volume Hamiltonian is
\(H_L=d\Gamma(h_L)\).  For finitely many one-particle modes with energies
\(\epsilon_1,\ldots,\epsilon_M\), the occupation-number basis gives
\[
\begin{aligned}
  \operatorname{Tr}_{\mathcal F_B}
  \exp\!\left[-\beta\,d\Gamma(h_M)\right]
  &=
  \sum_{k_1,\ldots,k_M\ge0}
  \exp\!\left[-\beta\sum_{j=1}^M k_j\epsilon_j\right]  \\
  &=
  \prod_{j=1}^M
  \frac{1}{1-\ee^{-\beta\epsilon_j}} .
\end{aligned}
\]
Passing to all torus modes is justified for \(\beta>0\) by convergence of
\[
  \log Z_B(\beta,L)
  =
  \sum_{n\in\mathbb Z^s}
  f(\beta\epsilon_n),
  \qquad
  f(x):=-\log(1-\ee^{-x}) .
\]
Indeed \(f(x)\le C\ee^{-x/2}\) for large \(x\), while the number of lattice
points with \(|n|_\infty=k\) is
\[
  N_s(k)=(2k+1)^s-(2k-1)^s
  \le 2s(2k+1)^{s-1}.
\]
For \(n\ne0\),
\[
  \epsilon_n\ge \frac{2\pi}{L}|n|_\infty .
\]
Set \(\tau=2\pi\beta/L\).  The remaining estimate is elementary but it is
important to separate the logarithmic singularity of \(f\) near the origin
from its exponential decay at infinity.  For \(0<x\le1\),
\[
  f(x)\le C(1+|\log x|),
\]
while for \(x\ge1\),
\[
  f(x)\le C\ee^{-x/2}.
\]
Therefore, for \(0<\tau\le1\),
\[
  \sum_{1\le k\le \tau^{-1}}(1+k)^{s-1}f(\tau k)
  \le
  C_s \tau^{-s}
  \int_0^1 u^{s-1}(1+|\log u|)\,\dd u ,
\]
and
\[
  \sum_{k> \tau^{-1}}(1+k)^{s-1}f(\tau k)
  \le
  C_s\sum_{k>\tau^{-1}} k^{s-1}\ee^{-\tau k/2}
  \le
  C'_s\tau^{-s}\int_1^\infty u^{s-1}\ee^{-u/2}\,\dd u .
\]
The two integrals are finite, so the nonzero-mode sum is bounded by
\(C_s\tau^{-s}\).  Hence
\[
  \log Z_B(\beta,L)
  \le
  f(\beta m)
  +
  C_s\left(\frac{L}{2\pi\beta}\right)^s .
\]
Since \(s\ge1\), the zero-mode term \(f(\beta m)\sim\log(\beta^{-1})\) is
absorbed into \(C_{s,m,L}\beta^{-s}\) on \(0<\beta\le1\).  Thus
\begin{equation}
  \log Z_B(\beta,L)
  \le
  C_{s,m,L}\,\beta^{-s},
  \qquad 0<\beta\le1 .
  \label{eq:free-fock-phase-space-growth}
\end{equation}
Equivalently,
\[
  Z_B(\beta,L)
  \le
  \exp(C_{s,m,L}\beta^{-s}) .
\]
The local nuclearity map replaces the global box trace by the localized map
\(A\mapsto \ee^{-\beta H}A\Omega\).  The theorem below is precisely the
operator-algebraic statement that a comparable phase-space estimate for
localized vectors is strong enough to create a split inclusion between
separated local algebras.  The free Fock calculation supplies the basic
mode-counting scale; the local theorem requires the sharper localization and
normality arguments described after the quoted theorem.

\begin{quotedtheorem}[Phase-space consequences of nuclearity]
\label{thm:nuclearity-split-type-iii-consequences}
\emph{Proof references.}\footnote{The phase-space condition and split
property are developed in D.~Buchholz and E.~H.~Wichmann,
``Causal independence and the energy-level density of states in local quantum
field theory,'' \emph{Communications in Mathematical Physics} \textbf{106}
(1986), 321--344, and in C.~D'Antoni and R.~Longo, ``Interpolation by type I
factors and the flip automorphism,'' \emph{Journal of Functional Analysis}
\textbf{51} (1983), 361--371.  Type \(\mathrm{III}_1\) results for local
algebras under standard relativistic regularity hypotheses are associated
with work of Fredenhagen, Buchholz, D'Antoni, Longo, and collaborators
\cite{BuchholzWichmann1986,DAntoniLongo1983}.}
Let a vacuum local net satisfy isotony, locality, covariance, positivity of
energy, uniqueness of the vacuum, Reeh--Schlieder for double cones, and the
Buchholz--Wichmann nuclearity condition.  Then every pair of double cones
\(\mathcal O_1\Subset\mathcal O_2\) is split:
\[
  \mathcal R(\mathcal O_1)
  \subset
  \mathcal N
  \subset
  \mathcal R(\mathcal O_2)
\]
for some type-I factor \(\mathcal N\).  With the usual additional regularity
assumptions used to exclude classical centers and finite local components,
the sharply local algebras are factors in the type-\(\mathrm{III}\) class; in
known models satisfying the full relativistic phase-space and scaling
regularity hypotheses, they are type \(\mathrm{III}_1\).
\end{quotedtheorem}

\paragraph{Functional-analytic mechanism of the split conclusion.}
The split part of the theorem converts an energy-damped compactness estimate
into a spatial tensor-product representation for separated algebras.  Let
\[
  \mathcal M_1=\mathcal R(\mathcal O_1),
  \qquad
  \mathcal M_2=\mathcal R(\mathcal O_2),
  \qquad
  \mathcal O_1\Subset\mathcal O_2 .
\]
Because \(\mathcal M_1\) and \(\mathcal M_2'\) commute, there is an algebraic
multiplication representation
\[
  m:\mathcal M_1\odot\mathcal M_2'\longrightarrow\mathcal B(\Hilb),
  \qquad
  m(A\otimes B')=AB' .
\]
If the inclusion is split, the normal tensor-product representation can be
seen directly.  A type-I factor
\(\mathcal N\) with
\(\mathcal M_1\subset\mathcal N\subset\mathcal M_2\) is spatial: there are
Hilbert spaces \(\mathcal K,\mathcal L\) and a unitary
\(U:\Hilb\to\mathcal K\otimes\mathcal L\) such that
\[
  U\mathcal N U^*=\mathcal B(\mathcal K)\otimes\mathbf 1 .
\]
Then
\[
  U\mathcal M_1U^*\subset\mathcal B(\mathcal K)\otimes\mathbf 1,
  \qquad
  U\mathcal M_2'U^*\subset U\mathcal N'U^*
  =
  \mathbf 1\otimes\mathcal B(\mathcal L).
\]
Therefore the product map is the restriction of the spatial normal
representation
\begin{equation}
  \mathcal M_1\bar\otimes\mathcal M_2'
  \longrightarrow
  \mathcal B(\mathcal K\otimes\mathcal L),
  \qquad
  A\otimes B'\longmapsto
  UAU^*\,UB'U^* .
  \label{eq:split-tensor-product-criterion}
\end{equation}
The converse implication, from normal extendability of the multiplication
representation to the existence of an interpolating type-I factor, is the
Doplicher--Longo split criterion.  It is part of the operator-algebraic input
in Quoted
Theorem~\ref{thm:nuclearity-split-type-iii-consequences}, not a consequence
of algebraic commutativity alone.  When the criterion applies, the unitary
identification can be chosen so that the two commuting algebras act as
\[
  \mathcal M_1\subset\mathcal B(\mathcal K)\otimes\mathbf 1,
  \qquad
  \mathcal M_2'\subset\mathbf 1\otimes\mathcal B(\mathcal L),
\]
and the interpolating type-I factor is
\[
  \mathcal N
  =
  U^{-1}\!\left(\mathcal B(\mathcal K)\otimes\mathbf 1\right)U .
\]
This criterion is the functional-analytic content behind the phrase
``approximate tensor factorization'': it is a normal tensor-product
representation for algebras separated by a positive collar.

\paragraph{Split states are positive-collar states.}
The split inclusion also produces a precise replacement for the informal
operation of assigning independent states to a region and its exterior.  Fix a
spatial implementation
\[
  U:\Hilb\to\mathcal K\otimes\mathcal L,
  \qquad
  U\mathcal N U^*=\mathcal B(\mathcal K)\otimes\mathbf 1 ,
\]
so that \(U\mathcal M_1U^*\subset\mathcal B(\mathcal K)\otimes\mathbf 1\)
and \(U\mathcal M_2'U^*\subset\mathbf 1\otimes\mathcal B(\mathcal L)\).  Let
\(\varphi_1\in(\mathcal M_1)_*^+\) and
\(\varphi_2\in(\mathcal M_2')_*^+\) be normal positive functionals.  Choose
normal positive extensions \(\widetilde\varphi_1\) and
\(\widetilde\varphi_2\) to \(\mathcal B(\mathcal K)\) and
\(\mathcal B(\mathcal L)\).  This is a pure von Neumann algebra extension
step: a normal positive functional on a von Neumann subalgebra of
\(\mathcal B(\mathcal K)\) has a normal positive extension to
\(\mathcal B(\mathcal K)\), with the same norm.  Their product normal
functional on
\(\mathcal B(\mathcal K)\bar\otimes\mathcal B(\mathcal L)\) restricts to a
normal positive functional on the von Neumann algebra generated by
\(\mathcal M_1\) and \(\mathcal M_2'\):
\begin{equation}
  \omega_{\varphi_1,\varphi_2}^{\mathcal O_1\Subset\mathcal O_2}(AB')
  =
  \varphi_1(A)\varphi_2(B'),
  \qquad
  A\in\mathcal M_1,\quad B'\in\mathcal M_2'.
  \label{eq:split-state-positive-collar-product}
\end{equation}
The right side shows independence of the chosen normal extensions on the
algebraic products \(AB'\), and normality gives the extension to the generated
von Neumann algebra.  Equation
\eqref{eq:split-state-positive-collar-product} is the exact mathematical
meaning of statistical independence at a positive separation.  The superscript
\(\mathcal O_1\Subset\mathcal O_2\) is part of the datum: the collar and the
interpolating type-I factor are not discarded.  If the collar is shrunk, the
theorem supplies no canonical limiting product state on
\(\mathcal R(\mathcal O)\vee\mathcal R(\mathcal O^\perp)\).  In a type-III
sharp local algebra there is no density matrix for \(\mathcal R(\mathcal O)\)
and no literal Hilbert-space tensor factor
\(\Hilb_{\mathcal O}\otimes\Hilb_{\mathcal O^\perp}\) from which such a state
could be obtained.  Split states are therefore regulator-independent local
QFT objects only after the positive collar has been specified.

The nuclearity hypothesis supplies the normality in
\eqref{eq:split-tensor-product-criterion}.  A nuclear decomposition of
\(\Theta_{\beta,\mathcal O}\) has the form
\[
  \ee^{-\beta H}A\vac
  =
  \sum_{n\ge1}\ell_n(A)\xi_n,
  \qquad
  \sum_n\|\ell_n\|\,\|\xi_n\|<\infty .
\]
Thus the set of vectors obtained by acting with the local unit ball and then
damping energy by \(\ee^{-\beta H}\) is contained in the absolutely convex
hull of a summable family.  This Banach-space statement is only the first
ingredient.  Since \(\mathcal M_1\) and \(\mathcal M_2'\) are von Neumann
algebras, the split criterion asks for normality, not only norm convergence.
Equivalently, for every vector pair \(\eta,\zeta\in\Hilb\), the functional
\[
  A\otimes B'
  \longmapsto
  \langle\eta,AB'\zeta\rangle
  \qquad
  (A\in\mathcal M_1,\ B'\in\mathcal M_2')
\]
must be represented, after the required limiting argument, by a normal
functional on \(\mathcal M_1\bar\otimes\mathcal M_2'\).  Finite sums of
separated normal functionals,
\[
  \sum_{j=1}^N\varphi_j(A)\psi_j(B'),
  \qquad
  \varphi_j\in(\mathcal M_1)_*,\quad
  \psi_j\in(\mathcal M_2')_*,
\]
are the elementary normal functionals on the spatial tensor product; a
summable infinite version is the analytic form in which the split proof
detects normal extendability of the multiplication representation.

The nuclear decomposition is used to produce such summable separated forms.
For a fixed \(B'\) and a fixed vector \(\eta\), the energy-damped matrix
coefficient
\[
  A\longmapsto
  \langle\eta,B'\ee^{-\beta H}A\vac\rangle
\]
has the absolutely convergent expansion
\[
  \sum_{n\ge1}\ell_n(A)\,\langle B'^*\eta,\xi_n\rangle .
\]
The displayed formula by itself does not prove splitness: the coefficient
functionals \(\ell_n\) in an arbitrary Banach-space nuclear decomposition need
not be normal, and the undamped vector functional
\((A,B')\mapsto\langle\eta,AB'\zeta\rangle\) is not obtained by algebraically
removing \(\ee^{-\beta H}\).  These two points are precisely where the local
QFT hypotheses enter.  The collar \(\mathcal O_1\Subset\mathcal O_2\) permits
small imaginary time translations whose tube domains do not cross the
spacelike complement of \(\mathcal O_2\).  Locality then moves operators in
\(\mathcal M_2'\) past the translated \(\mathcal M_1\)-operators, and the
spectral condition supplies the energy damping that makes the nuclear
expansion summable.  The remaining analytic estimates identify the resulting
bilinear vector functionals with limits, in the predual
\((\mathcal M_1\bar\otimes\mathcal M_2')_*\), of separated normal sums.  This
is the functional-analytic content hidden behind the short phrase
``nuclearity implies the split property.''

The finite-dimensional shadow is simple but useful for keeping the objects
straight.  If
\(\mathcal M_1\subset\mathcal B(\mathcal K)\otimes\mathbf 1\) and
\(\mathcal M_2'\subset\mathbf 1\otimes\mathcal B(\mathcal L)\), then any
density matrices \(r_1\) on \(\mathcal K\) and \(r_2\) on \(\mathcal L\) give
a normal product functional
\[
  A\otimes B'\longmapsto
  \operatorname{Tr}_{\mathcal K}(r_1A)\,
  \operatorname{Tr}_{\mathcal L}(r_2B')
  =
  \operatorname{Tr}_{\mathcal K\otimes\mathcal L}
  \!\left[(r_1\otimes r_2)(A\otimes B')\right].
\]
For a finite-rank map
\(\Theta(A)=\sum_j\ell_j(A)\xi_j\), every coefficient
\(\langle\eta,B'\Theta(A)\rangle\) is correspondingly a finite separated sum.
The infinite theorem is the statement that, under relativistic phase-space
and collar hypotheses, this finite separated normality survives the continuum
limit for the actual local multiplication representation.

This proof mechanism also explains the status of the constants in the
phase-space bound.  The exponential estimate
\[
  \|\Theta_{\beta,\mathcal O}\|_1
  \le
  \exp(C_{\mathcal O}\beta^{-s_{\mathcal O}})
\]
is stronger than mere nuclearity and is used to make the construction stable
as the collar, region, or thermodynamic comparison is varied.  The existence
of a single nuclear map for a fixed \((\beta,\mathcal O)\) gives compactness
at one scale; the uniform small-\(\beta\) bound is the continuum
phase-space statement needed in applications.

\paragraph{Type classification is a separate sharp-region theorem.}
The split theorem and the type-classification theorem concern different
objects.  Let \(\mathcal M\) be a von Neumann algebra.  Its center is
\[
  Z(\mathcal M)=\mathcal M\cap\mathcal M',
\]
and \(\mathcal M\) is a factor when \(Z(\mathcal M)=\mathbb C\mathbf 1\).  A
projection \(p\in\mathcal M\) is minimal when
\[
  p\mathcal M p=\mathbb C p .
\]
A type-I factor has minimal projections and is isomorphic to
\(\mathcal B(\mathcal K)\) for a Hilbert space \(\mathcal K\).  A factor is
type III when it has no nonzero finite projections, equivalently when it has no
nonzero normal semifinite trace.  For a type-III factor the finer Connes
invariant may be described by
\[
  S(\mathcal M)
  =
  \bigcap_{\varphi}
  \operatorname{Spec}(\Delta_\varphi),
\]
where the intersection is over faithful normal semifinite weights
\(\varphi\) and \(\Delta_\varphi\) is the corresponding modular operator in
the standard form.  The type \(\mathrm{III}_1\) case is characterized by
\[
  S(\mathcal M)=\mathbb R_{\ge0}.
\]

These definitions show exactly why the type statement cannot be read off from
the split inclusion alone.  The split inclusion supplies, for
\(\mathcal O_1\Subset\mathcal O_2\), an auxiliary type-I factor
\(\mathcal N\) between
\(\mathcal R(\mathcal O_1)\) and \(\mathcal R(\mathcal O_2)\).  It does not
say that either endpoint algebra is type I.  Conversely, a finite-regulator
lattice system has type-I local algebras and may satisfy an exact tensor
factorization at positive lattice separation; that finite statement is not the
continuum sharp-region conclusion.  The theorem-boundary problem for local QFT
is to prove, in the continuum representation, the factor property, absence of
finite local projections, and the modular spectrum needed for
type-\(\mathrm{III}_1\).  Those inputs use locality, covariance, the spectrum
condition, Reeh--Schlieder standardness, additivity or scaling regularity, and
phase-space control in a way that is not captured by the existence of a
positive-collar type-I interpolant.

The quoted theorem has two distinct messages.  The split inclusion says that a
positive collar between regions allows a type-I interpolant and hence an
approximate tensor product.  The type-\(\mathrm{III}\) statement says that the
collar cannot be shrunk to zero without losing the type-I description.  In
particular, a formal notation such as
\[
  \Hilb\stackrel{\rm formal}{\neq}
  \Hilb_{\mathcal O}\otimes\Hilb_{\mathcal O^\perp}
\]
is not a technical inconvenience; it is the local algebraic structure of a
continuum relativistic QFT.  Local entropy therefore has to be understood
through split limits, relative entropy, or other regulator-dependent
procedures whose limiting content is stated explicitly.

The type-\(\mathrm{III}_1\) identification is a further sharp-local-algebra
statement.  Its input is not the split property alone but the combination of
relativistic locality, vacuum cyclicity and separatingness, additivity or
scaling regularity, and the absence of finite-dimensional local summands.
Under those hypotheses the local algebra has no minimal projections, no
faithful normal trace, and no type-I density-matrix description.  The split
property supplies type-I factors only at positive separation; the sharp
region algebra belongs to the type-III class.

\section{Modular Hamiltonians and Energy Bounds}
\label{sec:modular-hamiltonian-relative-entropy-energy-bounds}

Let \(\omega\) be a faithful normal state on \(\mathcal R\), represented by a
cyclic separating vector \(\Omega\).  The modular Hamiltonian of the pair
\((\mathcal R,\omega)\) is
\[
  K_\omega=-\log\Delta_\omega .
\]
It is an operator on the GNS Hilbert space of the algebra-state pair.  When
\(\mathcal R=\mathcal B(\mathcal K_A)\otimes 1\) in a finite bipartite
system and \(\omega\) has reduced density matrix \(\rho_A^\omega\), the
restriction of \(K_\omega\) to the left standard representation is
\(-\log\rho_A^\omega\) up to the right-action term that cancels in left
expectation differences.  Thus the finite-dimensional identity
\[
  S(\rho_A^\psi\|\rho_A^\omega)
  =
  \operatorname{Tr}\rho_A^\psi
    (\log\rho_A^\psi-\log\rho_A^\omega)
\]
can be rewritten as
\begin{equation}
  S(\rho_A^\psi\|\rho_A^\omega)
  =
  \Delta_\psi\langle K_A^\omega\rangle
  -
  \Delta_\psi S_A,
  \qquad
  K_A^\omega=-\log\rho_A^\omega .
  \label{eq:finite-relative-entropy-first-law-form}
\end{equation}
Here
\[
  \Delta_\psi\langle K_A^\omega\rangle
  =
  \operatorname{Tr}(\rho_A^\psi-\rho_A^\omega)K_A^\omega,
  \qquad
  \Delta_\psi S_A
  =
  -\operatorname{Tr}\rho_A^\psi\log\rho_A^\psi
  +\operatorname{Tr}\rho_A^\omega\log\rho_A^\omega .
\]
Equation \eqref{eq:finite-relative-entropy-first-law-form} is often called
the entropy--modular-energy identity.  In a type-\(\mathrm{III}\) local
algebra the two terms on the right are not separately defined without a
chosen split inclusion or regulator, but the left side is the Araki relative
entropy \eqref{eq:araki-relative-entropy}.  The mathematically primary object
is therefore
\[
  S_{\mathcal R}(\psi\|\omega),
\]
not an unqualified local von Neumann entropy.

\paragraph{Finite-dimensional first variation of relative entropy.}
Let \(\rho_\lambda=\rho_0+\lambda\dot\rho+O(\lambda^2)\) be a differentiable
family of faithful density matrices with \(\operatorname{Tr}\dot\rho=0\).
Let \(K_0=-\log\rho_0\).  Then
\[
  \left.\frac{\dd}{\dd\lambda}\right|_{\lambda=0}
  S(\rho_\lambda\|\rho_0)=0
\]
is equivalent to
\begin{equation}
  \left.\frac{\dd}{\dd\lambda}\right|_{\lambda=0}
  S_{\rm vN}(\rho_\lambda)
  =
  \operatorname{Tr}(\dot\rho K_0).
  \label{eq:first-law-entanglement}
\end{equation}
The calculation uses
\[
  S(\rho_\lambda\|\rho_0)
  =
  \operatorname{Tr}\rho_\lambda\log\rho_\lambda
  -
  \operatorname{Tr}\rho_\lambda\log\rho_0 .
\]
The Fr\'echet derivative of \(X\mapsto\operatorname{Tr}X\log X\) at a faithful
matrix \(\rho_0\) is
\[
  \operatorname{Tr}\dot\rho(\log\rho_0+1).
\]
The \(1\)-term vanishes because \(\operatorname{Tr}\dot\rho=0\).  Therefore
\[
  \left.\frac{\dd}{\dd\lambda}\right|_{0}
  S(\rho_\lambda\|\rho_0)
  =
  \operatorname{Tr}\dot\rho\log\rho_0
  -
  \operatorname{Tr}\dot\rho\log\rho_0
  =
  0.
\]
On the other hand,
\[
  \left.\frac{\dd}{\dd\lambda}\right|_{0}
  S_{\rm vN}(\rho_\lambda)
  =
  -\operatorname{Tr}\dot\rho\log\rho_0
  =
  \operatorname{Tr}\dot\rho K_0,
\]
which proves \eqref{eq:first-law-entanglement}.  The same identity in local
QFT is obtained by applying this finite calculation to split inclusions and
then passing only those quantities with regulator-independent limits to the
sharp algebra.

The inequality
\[
  S_{\mathcal R}(\psi\|\omega)\ge0
\]
is the modular-theoretic source of many energy bounds.  In a regulated
type-I approximation it gives
\[
  \Delta_\psi S_A\le \Delta_\psi\langle K_A^\omega\rangle .
\]
When \(\mathcal R\) is a wedge algebra and \(\omega\) is the vacuum,
Bisognano--Wichmann identifies
\[
  K_{W_R,\omega}=2\pi K_R,
  \qquad U(\Lambda_R(s))=\ee^{\ii sK_R},
\]
because \(\Delta_{W_R}^{\ii t}=U(\Lambda_R(-2\pi t))\).  If the theory has a
stress tensor normalized so that \(K_R\) is the boost charge, then formally on
the \(x^0=0\) Cauchy surface
\begin{equation}
  K_{W_R,\omega}
  =
  2\pi\int_{x^1>0} x^1\,T_{00}(0,\mathbf x)\,\dd^{D-1}\mathbf x .
  \label{eq:wedge-modular-hamiltonian-stress-tensor}
\end{equation}
Equation \eqref{eq:wedge-modular-hamiltonian-stress-tensor} is not a
definition of the local algebra; it is the consequence of the geometric
modular-flow theorem plus the independently normalized stress tensor.  The
standard derivations of the Bekenstein-type bound for half-spaces and of the
averaged null energy condition use three inputs: positivity of Araki relative
entropy, monotonicity of relative entropy under inclusion of von Neumann
algebras, and the geometric form of the vacuum modular Hamiltonian for wedges
or balls.  Whenever one of these inputs is absent, the claimed bound must be
restated as a hypothesis or as a regulated theorem.

\section{Half-Sided Modular Inclusions}
\label{sec:half-sided-modular-inclusions}

Modular theory can also generate spacetime symmetry from inclusions of
algebras.  This is the mechanism behind the use of modular inclusions in
chiral CFT, wedge-local constructions, and several reconstruction theorems.

\begin{definition}[Half-sided modular inclusion]
\label{def:half-sided-modular-inclusion}
Let \(\mathcal N\subset\mathcal M\) be von Neumann algebras on \(\Hilb\), and
let \(\Omega\) be cyclic and separating for both.  Let
\(\sigma_t^{\mathcal M}\) be the modular automorphism group of
\((\mathcal M,\Omega)\).  The inclusion is right half-sided modular if
\[
  \sigma_t^{\mathcal M}(\mathcal N)\subset\mathcal N,
  \qquad t\ge0,
\]
and left half-sided modular if the same inclusion holds for \(t\le0\).
\end{definition}

\begin{quotedtheorem}[Borchers--Wiesbrock structure]
\label{thm:borchers-wiesbrock-half-sided-inclusion}
\emph{Proof references.}\footnote{Borchers' commutation relations and
Wiesbrock's reconstruction theorem are proved in H.-J.~Borchers,
``The CPT theorem in two-dimensional theories of local observables,''
\emph{Communications in Mathematical Physics} \textbf{143} (1992),
315--332, and H.-W.~Wiesbrock, ``Half-sided modular inclusions of von
Neumann algebras,'' \emph{Communications in Mathematical Physics}
\textbf{157} (1993), 83--92.  See also Guido--Longo--Wiesbrock for the
connection with conformal nets \cite{Borchers1992,Wiesbrock1993,Borchers2000}.}
Let \(\mathcal N\subset\mathcal M\) be a standard right half-sided modular
inclusion, meaning here that \(\Omega\) is cyclic and separating for
\(\mathcal M\), for \(\mathcal N\), and for the relative commutant
\(\mathcal N'\cap\mathcal M\), and that
\[
  \sigma_t^{\mathcal M}(\mathcal N)\subset\mathcal N,
  \qquad t\ge0 .
\]
Then there is a strongly continuous one-parameter unitary group.  With the
Stone sign convention used below for physical inward light-ray translations,
write it as
\[
  T(a)=\ee^{-\ii aP},\qquad P\ge0,
\]
such that \(T(a)\Omega=\Omega\), \(T(a)\mathcal M T(a)^*\subset\mathcal M\)
for \(a\ge0\), and the modular group of \(\mathcal M\) acts by dilations:
\begin{equation}
  \Delta_{\mathcal M}^{\ii t}T(a)\Delta_{\mathcal M}^{-\ii t}
  =
  T(\ee^{+2\pi t}a).
  \label{eq:borchers-commutation-relation}
\end{equation}
Equivalently, with \(K_{\mathcal M}=-\log\Delta_{\mathcal M}\),
\[
  \Delta_{\mathcal M}^{\ii t}=\ee^{-\ii tK_{\mathcal M}},
  \qquad
  [K_{\mathcal M},P]=+2\pi\ii P
\]
on the common invariant core of analytic vectors.
For a left half-sided modular inclusion the same statement is obtained after
reversing the modular parameter; equivalently the dilation factor in
\eqref{eq:borchers-commutation-relation} is replaced by
\(\ee^{-2\pi t}a\), and the infinitesimal commutator has the opposite sign.
\end{quotedtheorem}

\paragraph{The modular-inclusion mechanism.}
The theorem is not the differentiated commutator below.  Its content is the
construction of the translation unitary \(T(a)\) and the positivity of its
generator from inclusion data alone.  The standardness assumption provides
three standard pairs:
\[
  (\mathcal M,\Omega),\qquad
  (\mathcal N,\Omega),\qquad
  (\mathcal N'\cap\mathcal M,\Omega).
\]
Equivalently, in standard-subspace language, the closed real subspaces
\[
  H_{\mathcal M}
  =
  \overline{\{A\Omega:\ A=A^*\in\mathcal M\}}^{\,\Hilb},
  \qquad
  H_{\mathcal N}
  =
  \overline{\{A\Omega:\ A=A^*\in\mathcal N\}}^{\,\Hilb}
\]
are standard real subspaces: \(H\cap\ii H=\{0\}\) and
\(H+\ii H\) is dense.  Their Tomita operators
\[
  s_{\mathcal M}(h+\ii k)=h-\ii k,\qquad h,k\in H_{\mathcal M},
\]
and similarly for \(H_{\mathcal N}\), have polar decompositions with the same
modular objects as the von Neumann algebras.  The half-sided condition
\(\Delta_{\mathcal M}^{\ii t}H_{\mathcal N}\subset H_{\mathcal N}\) for
\(t\ge0\) is therefore a geometric statement about the position of two
standard real subspaces under the modular dilations of the larger one.

The finite-dimensional comparison is rigid.  Suppose, for comparison, that
\(\mathcal M\) is finite
dimensional and that
\(\sigma_t^{\mathcal M}(\mathcal N)\subset\mathcal N\) for all \(t\ge0\).
The automorphism \(\sigma_t^{\mathcal M}\) is a linear isomorphism of the
finite-dimensional vector space \(\mathcal M\).  Its restriction maps the
finite-dimensional subspace \(\mathcal N\) injectively into itself, and hence
onto itself:
\[
  \sigma_t^{\mathcal M}(\mathcal N)=\mathcal N,\qquad t\ge0 .
\]
Applying the same equality to \(-t\) by the group law gives invariance for all
\(t\in\mathbb R\).  Thus the half-sided condition becomes ordinary modular
invariance in finite dimension.  The strict inclusion phenomenon in Quoted
Theorem~\ref{thm:borchers-wiesbrock-half-sided-inclusion} depends on the
infinite-dimensional standard-subspace geometry of local algebras: an
injective continuous motion can move a closed real subspace properly inside
itself without contradicting finite-dimensional rank.

Borchers--Wiesbrock theory proves that such a pair of standard real subspaces
carries a positive-energy representation of the affine group of the line.  In
the Stone sign convention of the theorem statement, this representation is
generated by the modular dilation group of \(\mathcal M\) and a translation
group \(T(a)=\ee^{-\ii aP}\), \(P\ge0\), satisfying
\[
  \Delta_{\mathcal M}^{\ii t}T(a)\Delta_{\mathcal M}^{-\ii t}
  =
  T(\ee^{+2\pi t}a).
\]
The positivity \(P\ge0\) is the analytic half-sidedness input: it is the
operator-algebraic analogue of the spectrum condition for the translation
reconstructed from the inclusion.  Once the unitary group has been obtained
on the standard subspaces, standardness of the relative commutant lifts the
standard-subspace inclusion back to the von Neumann algebra inclusion and
gives
\[
  T(a)\mathcal M T(a)^*\subset\mathcal M,\qquad a\ge0 .
\]
Thus the theorem reconstructs a translation-dilation semigroup from modular
data; the Lie-algebra commutator is only the infinitesimal shadow of this
stronger statement.

\paragraph{Derivation of the displayed commutator.}
The existence of \(T(a)\) and the positivity \(P\ge0\) require the full
half-sided modular inclusion result, Quoted
Theorem~\ref{thm:borchers-wiesbrock-half-sided-inclusion}.  Once
\eqref{eq:borchers-commutation-relation} is known, the
commutator follows by differentiating twice.  For the right half-sided
convention displayed in \eqref{eq:borchers-commutation-relation}, first
differentiate in \(a\) at \(a=0\):
\[
  \Delta_{\mathcal M}^{\ii t}P\Delta_{\mathcal M}^{-\ii t}
  =
  \ee^{+2\pi t}P .
\]
Since \(\Delta_{\mathcal M}^{\ii t}=\ee^{-\ii tK_{\mathcal M}}\), the
derivative at \(t=0\) gives
\[
  -\ii[K_{\mathcal M},P]=+2\pi P,
\]
and hence \([K_{\mathcal M},P]=+2\pi\ii P\).  In the opposite half-sided
convention the Borchers relation has \(\ee^{-2\pi t}a\) in place of
\(\ee^{+2\pi t}a\), so the differentiated commutator carries the opposite
sign.

For the standard right wedge, take \(\mathcal M=\mathcal R(W_R)\) and let
\(\mathcal N\) be the algebra of the wedge translated inward along a lightlike
direction.  The inclusion is half-sided because the Lorentz boost rescales
that lightlike translation without leaving the wedge family.  The sign is
convention-sensitive, and the convention used in this chapter is the following
one.

\begin{example}[Right-wedge light-ray inclusion and the half-sided sign]
\label{ex:right-wedge-half-sided-sign}
Let
\[
  e_+=(1,1,0,\ldots,0),
  \qquad
  W_R(a)=W_R+a e_+,\qquad a>0 .
\]
Then \(W_R(a)\subset W_R\).  Indeed, if
\((x^0,x^1)\in W_R\) and \(y=x+a e_+\), then
\[
  y^1-y^0=x^1-x^0>0,\qquad
  y^1+y^0=x^1+x^0+2a>0,
\]
so \(y^1>|y^0|\).  Moreover
\begin{equation}
  b\ge a\quad\Longrightarrow\quad W_R(b)\subset W_R(a),
  \label{eq:right-wedge-light-ray-nesting}
\end{equation}
because translating farther along \(e_+\) moves the wedge deeper inside
itself.  Assume the Bisognano--Wichmann identification in the sign convention
used below,
\[
  \sigma_t^{W_R}
  =
  \operatorname{Ad}\Delta_{W_R}^{\ii t}
  =
  \operatorname{Ad}U(\Lambda_R(-2\pi t)).
\]
Since \(\Lambda_R(s)e_+=\ee^s e_+\), isotony and covariance give
\begin{equation}
  \sigma_t^{W_R}\bigl(\mathcal R(W_R(a))\bigr)
  =
  \mathcal R\bigl(W_R(\ee^{-2\pi t}a)\bigr).
  \label{eq:right-wedge-half-sided-scaling}
\end{equation}
Combining \eqref{eq:right-wedge-light-ray-nesting} and
\eqref{eq:right-wedge-half-sided-scaling}, this translated-wedge inclusion is
left half-sided in Definition~\ref{def:half-sided-modular-inclusion}:
\[
  \sigma_t^{W_R}\bigl(\mathcal R(W_R(a))\bigr)
  \subset
  \mathcal R(W_R(a))
  \quad\Longleftrightarrow\quad
  t\le0 .
\]
Changing the modular-flow sign convention, or using the corresponding
oppositely directed light-ray inclusion, changes the word ``left'' to
``right''.  The calculation verifies only the geometric hypothesis in this
example.  The Borchers--Wiesbrock theorem supplies the converse structural
content: from the half-sided modular inclusion it reconstructs the
translation unitary and its positive generator.
\end{example}

\paragraph{Physical light-ray translations in a covariant net.}
If the net already carries a Poincare representation, the abstract affine
translation in the preceding paragraph has a visible candidate.  With the
translation convention of
Definition~\ref{def:poincare-covariant-vacuum-hilbert-datum},
\[
  U(x,\mathbf 1)
  =
  \int_{\overline V_+}\ee^{\ii x^\mu p_\mu}\,\dd E(p),
  \qquad p_\mu=\eta_{\mu\nu}p^\nu ,
\]
define the positive light-ray momentum
\[
  P_+^{\rm phys}:=P^0-P^1
  =
  \int_{\overline V_+}(p^0-p^1)\,\dd E(p).
\]
The spectrum condition gives \(P_+^{\rm phys}\ge0\), since
\(p^0\ge |p^1|\) on \(\overline V_+\).  For
\(e_+=(1,1,0,\ldots,0)\), however,
\[
  e_+^\mu p_\mu=-p^0+p^1=-(p^0-p^1),
\]
so the physical inward light-ray translation is
\[
  U(ae_+,\mathbf 1)=\ee^{-\ii aP_+^{\rm phys}},\qquad a\ge0 .
\]
This is the reason for the Stone sign convention in
Quoted Theorem~\ref{thm:borchers-wiesbrock-half-sided-inclusion}.  Isotony
and covariance give
\[
  U(ae_+,\mathbf 1)\mathcal R(W_R)U(ae_+,\mathbf 1)^*
  =
  \mathcal R(W_R(a))
  \subset
  \mathcal R(W_R),
  \qquad a\ge0,
\]
and the Bisognano--Wichmann modular action gives the affine covariance
\[
  \Delta_{W_R}^{\ii t}U(ae_+,\mathbf 1)\Delta_{W_R}^{-\ii t}
  =
  U(\Lambda_R(-2\pi t)ae_+,\mathbf 1)
  =
  U(\ee^{-2\pi t}ae_+,\mathbf 1).
\]
Thus, in a Poincare-covariant net satisfying Bisognano--Wichmann, the
physical light-ray translations realize the left half-sided version of the
positive-energy affine representation reconstructed by the
Borchers--Wiesbrock theorem.  The theorem's converse content is not this sign
calculation: it is that a standard half-sided modular inclusion by itself
already forces such a positive translation-dilation representation.

Thus the algebraic inclusion knows the translation-dilation subgroup before
one has written down a point field.  This is one reason modular theory is not
only a structural theorem: it reconstructs geometric motion from the order
structure of local algebras.

Let
\[
  W_R=\{x\in\mathbb M^D:\ x^1>|x^0|\}
\]
be the standard right wedge, and let \(\Lambda_R(s)\) be the boost in the
\((x^0,x^1)\)-plane,
\[
  \Lambda_R(s):
  \begin{pmatrix}x^0\\ x^1\end{pmatrix}
  \longmapsto
  \begin{pmatrix}
    \cosh s & \sinh s\\
    \sinh s & \cosh s
  \end{pmatrix}
  \begin{pmatrix}x^0\\ x^1\end{pmatrix}.
\]
\begin{remark}[Bisognano--Wichmann theorem boundary for wedges]
\label{rem:bisognano-wichmann-wedges-boundary}
Let a Poincare-covariant quantum field theory on Minkowski space be generated
by Wightman fields satisfying the Wightman domain, covariance, locality,
vacuum, and spectrum hypotheses.  Let
\[
  \mathcal R(W)
\]
be the von Neumann algebra generated by fields smeared in a wedge \(W\), and
let \(\vac\) be the unique invariant vacuum.  Then \(\vac\) is cyclic and
separating for every wedge algebra.  For the standard right wedge \(W_R\), let
\((J_{W_R},\Delta_{W_R})\) be the Tomita--Takesaki modular data of
\((\mathcal R(W_R),\vac)\).  Then
\[
  \Delta_{W_R}^{\ii t}=U(\Lambda_R(-2\pi t)),
  \qquad t\in\mathbb R.
\]
If \(W=gW_R\), then the modular group of \((\mathcal R(W),\vac)\) is
\[
  \Delta_W^{\ii t}
  =
  U\!\left(g\Lambda_R(-2\pi t)g^{-1}\right),
  \qquad t\in\mathbb R.
\]
Moreover the modular conjugation implements the corresponding wedge reflection
\(j_W\) on the local net:
\[
  J_W\mathcal R(\mathcal O)J_W=\mathcal R(j_W\mathcal O)
\]
for every local region \(\mathcal O\) in the indexing class.
On a charged field algebra reconstructed from the observable net, the same
antiunitary operation includes the charge-conjugation action on charged field
coordinates.
This is a theorem boundary in the present monograph, with its scope recorded
explicitly.  The proof in the Wightman setting has two logically separate
components.  The first component is the polynomial-core KMS strip
proved in Theorem~\ref{thm:wightman-polynomial-wedge-kms} below from the
spectrum condition, complex boosts, and wedge locality.  The second component
is the operator-domain closure theorem that identifies the closed Tomita
operator of \((\mathcal R(W_R),\vac)\), including the modular conjugation and
the passage from field-polynomial boundary values to the full wedge von
Neumann algebra.  That closure theorem is the part supplied by the original
Bisognano--Wichmann work and by the AQFT treatments of modular covariance.
\end{remark}

\begin{theorem}[Wightman polynomial-core wedge KMS mechanism]
\label{thm:wightman-polynomial-wedge-kms}
Let \(\Phi_a\) be Wightman fields on the common invariant domain
\(\mathcal D\), transforming covariantly under the proper orthochronous
Poincare group and satisfying the spectrum condition and graded locality.  Let
\(\mathfrak P(W_R)\) be the unital \(*\)-algebra generated on \(\mathcal D\)
by smeared fields \(\Phi_a(f)\) with \(\operatorname{supp}f\subset W_R\), and
let
\[
  \alpha_s^R(A)=U(\Lambda_R(s))AU(\Lambda_R(s))^{-1}.
\]
For
\[
  A=\Phi_{a_1}(f_1)\cdots\Phi_{a_m}(f_m),
  \qquad
  B=\Phi_{b_1}(g_1)\cdots\Phi_{b_n}(g_n),
  \qquad
  \operatorname{supp}f_i,\operatorname{supp}g_j\subset W_R .
\]
there is a function \(F_{A,B}\), holomorphic for \(0<\operatorname{Im}z<2\pi\)
and with tempered boundary values on the two horizontal edges, such that
\begin{equation}
  F_{A,B}(s)=\langle\vac,A\alpha_s^R(B)\vac\rangle,
  \qquad
  F_{A,B}(s+2\pi\ii)=\langle\vac,\alpha_s^R(B)A\vac\rangle .
  \label{eq:wightman-polynomial-core-kms}
\end{equation}
Thus the vacuum functional satisfies the boost \(2\pi\)-KMS boundary relation
on the Wightman polynomial core of the wedge.  For a fermionic field algebra
the statement is the graded KMS relation on field coordinates; after
restriction to the observable algebra the grading signs cancel.
\end{theorem}

\begin{proof}
The theorem is an analytic-domain statement: the boost generator supplies the
real flow, while the KMS boundary relation requires holomorphic continuation
and locality.  The initial Tomita operator on polynomial wedge vectors is
\[
  S_{W_R,0}A\vac=A^*\vac,
  \qquad A\in\mathfrak P(W_R).
\]
Reeh--Schlieder gives density of \(\mathfrak P(W_R)\vac\), and
Corollary~\ref{cor:reeh-schlieder-separatingness} supplies separatingness of
\(\vac\) for \(\mathcal R(W_R)\) from locality and cyclicity in the opposite
wedge.  These are the standardness inputs that make the initial Tomita
operator closable in the full Bisognano--Wichmann theorem.  The polynomial
KMS strip itself is
obtained directly from Wightman analyticity.  Suppressing finite spin matrices
in the covariance law, define the smeared boundary value
\begin{equation}
\begin{aligned}
  F_{A,B}(z)
  &=
  \int
  \prod_i \dd x_i\, f_i(x_i)
  \prod_j \dd y_j\, g_j(y_j)                                      \\
  &\quad\times
  W_{a_1\ldots a_m b_1\ldots b_n}
  \bigl(x_1,\ldots,x_m,\Lambda_R(z)y_1,\ldots,\Lambda_R(z)y_n\bigr),
\end{aligned}
  \label{eq:bw-wightman-strip-function}
\end{equation}
with the usual component matrices inserted for non-scalar fields.  The
spectrum condition supplies the tube holomorphy as follows.  After rewriting
an ordered Wightman distribution in relative variables, its Fourier transform
is supported where every partial momentum sum lies in the closed future cone.
If a relative coordinate is shifted as \(\xi\mapsto\xi-\ii\eta\) with
\(\eta\in V_+\), the Fourier factor
\(\exp(-\ii p\cdot(\xi-\ii\eta))\) contains \(\exp(p\cdot\eta)\); in the
mostly-plus convention \(p\cdot\eta\le0\) for
\(p,\eta\in\overline V_+\).  Thus the spectral support gives exponential
damping in the tube selected by the ordering, and the Wightman distribution is
recovered as a boundary value.  For
\(z=s+\ii u\) and \(0<u<\pi\), the imaginary part of
\(\Lambda_R(z)y\), \(y\in W_R\), lies in the future timelike cone in the
mostly-plus convention.  Therefore the differences between each earlier real
argument \(x_i\) and each complex-boosted later argument have imaginary part
in the past tube, so the ordered distribution in
\eqref{eq:bw-wightman-strip-function} is holomorphic in the first half-strip
and has tempered boundary values.

At \(u=\pi\), \(\Lambda_R(s+\ii\pi)W_R=W_L\), the left wedge.  Points of
\(W_L\) are spacelike to points of \(W_R\).  Graded locality therefore permits
the boosted fields in \(B\) to pass through the fields in \(A\), with the sign
determined by the field grading.  On the observable subalgebra this sign is
one.  After this reordering, the same tube analyticity gives the continuation
through the second half-strip \(\pi<u<2\pi\).  The lower and upper boundary
values are consequently
\begin{equation}
  F_{A,B}(s+2\pi\ii)
  =
  \omega\!\left(\alpha_s^R(B)A\right),
  \qquad
  F_{A,B}(s)=\omega\!\left(A\alpha_s^R(B)\right),
  \label{eq:bw-wightman-kms-boundary}
\end{equation}
where \(\omega(C)=\langle\vac,C\vac\rangle\).  This is
\eqref{eq:wightman-polynomial-core-kms}.  The full modular theorem adds the
closure statement
\[
  \overline{S_{W_R,0}}=J_{W_R}\Delta_{W_R}^{1/2}.
\]
The operator \(\Delta_{W_R}^{1/2}\) is then obtained by analytic continuation
of boosts to imaginary value \(\ii\pi\), and the remaining antiunitary maps the
wedge to its geometric reflection.  That closed-operator identification is the
additional domain theorem named in
Remark~\ref{rem:bisognano-wichmann-wedges-boundary}.
\end{proof}

\begin{corollary}[PCT from wedge modular covariance]
\label{cor:aqft-pct-from-modular-covariance}
Let \(D=4\).  Let a vacuum representation of a local observable net satisfy
isotony, locality, Poincare covariance, the spectrum condition, uniqueness of
the invariant vacuum, wedge cyclicity and separatingness, wedge duality, and
the Bisognano--Wichmann modular-covariance conclusion
\[
  \Delta_{W_R}^{\ii t}=U(\Lambda_R(-2\pi t)),
  \qquad
  J_{W_R}\mathcal R(\mathcal O)J_{W_R}=\mathcal R(j_R\mathcal O)
\]
for the standard right wedge.  Let \(R_\perp(\pi)\) be the spatial rotation by
\(\pi\) in the \((x^2,x^3)\)-plane, so that
\[
  R_\perp(\pi)j_R x=-x .
\]
Then
\begin{equation}
  \Theta_{\rm PCT}:=U(R_\perp(\pi))J_{W_R}
  \label{eq:aqft-pct-modular-conjugation}
\end{equation}
is an antiunitary operator satisfying
\[
  \Theta_{\rm PCT}\vac=\vac,
  \qquad
  \Theta_{\rm PCT}U(a,\Lambda)\Theta_{\rm PCT}^{-1}=U(-a,\Lambda),
\]
and
\begin{equation}
  \Theta_{\rm PCT}\mathcal R(\mathcal O)\Theta_{\rm PCT}^{-1}
  =
  \mathcal R(-\mathcal O).
  \label{eq:aqft-pct-net-action}
\end{equation}
For a Doplicher--Roberts charged field algebra reconstructed from the
observable category, the induced antiunitary sends each charged field
multiplet to the conjugate charged multiplet, in agreement with the Wightman
PCT theorem of Theorem~\ref{thm:wightman-pct}.
\end{corollary}

\begin{proof}
The antiunitarity of \(\Theta_{\rm PCT}\) follows from antiunitarity of the
modular conjugation \(J_{W_R}\) and unitarity of \(U(R_\perp(\pi))\).  The
displayed hypotheses identify \(J_{W_R}\) with the geometric wedge reflection
\(j_R\) on the net.  Conjugating afterward by the proper spatial
rotation \(R_\perp(\pi)\) changes the geometric action from \(j_R\) to the
total inversion \(x\mapsto -x\), proving
\eqref{eq:aqft-pct-net-action}.  Both \(J_{W_R}\) and the rotation preserve
the vacuum, so \(\Theta_{\rm PCT}\vac=\vac\).

The covariance identity follows by applying the same geometric conjugation to
translations and proper Lorentz transformations.  Since the total inversion
commutes with every proper Lorentz transformation and sends \(a\) to \(-a\),
the automorphism of the Poincare group is
\[
  (a,\Lambda)\longmapsto(-a,\Lambda).
\]
Antiunitarity then gives the same translation-generator convention as in
\eqref{eq:pct-poincare-implementation}.  Finally, in the
Doplicher--Roberts reconstruction the modular conjugation maps a localized
endomorphism to its conjugate localized endomorphism; on the reconstructed
field algebra this is exactly the passage from a charged multiplet to the
conjugate charged multiplet.
\end{proof}

The cyclic and separating hypotheses needed for the modular data of wedge
algebras are supplied in the Wightman setting by
Theorem~\ref{thm:reeh-schlieder-weakly-additive-net} and
Corollary~\ref{cor:reeh-schlieder-separatingness}: cyclicity is the
positive-energy local-density theorem, while separatingness for a wedge uses
locality with the opposite wedge.  In a purely algebraic formulation not
assumed to arise from Wightman fields, the conclusion of
Remark~\ref{rem:bisognano-wichmann-wedges-boundary} is an explicit
modular-covariance hypothesis on the vacuum net unless it has been proved from
the chosen AQFT assumptions.  Under that theorem or hypothesis, modular flow
is geometric Lorentz boost flow.  The result connects locality, positivity of
energy, the KMS structure of Definition~\ref{def:kms-state} for accelerated
observers, and the PCT antiunitary stated in
Corollary~\ref{cor:aqft-pct-from-modular-covariance}.

Indeed, define the boost automorphism of the standard wedge algebra by
\[
  \alpha_s^R(A)=
  U(\Lambda_R(s))AU(\Lambda_R(s))^{-1},
  \qquad A\in\mathcal R(W_R).
\]
Bisognano--Wichmann says \(\sigma_t=\alpha_{-2\pi t}^R\), so the modular KMS
property implies that \(\omega(A)=\langle\vac,A\vac\rangle\), restricted to
\(\mathcal R(W_R)\), is a \(2\pi\)-KMS state for the boost dynamics
\(\alpha_s^R\):
\[
  F_{A,B}(s)=\omega(A\alpha_s^R(B)),
  \qquad
  F_{A,B}(s+2\pi\ii)=\omega(\alpha_s^R(B)A).
\]
Thus the vacuum is thermal with respect to the boost generator, even though it
is a pure state on the global algebra.  Along a uniformly accelerated orbit
whose proper time is related to boost parameter by \(s=a\tau\), this becomes
the Unruh inverse temperature \(2\pi/a\) in units \(\hbar=c=k_B=1\).

Figure~\ref{fig:modular-wedge-flow} shows the Bisognano--Wichmann wedge
geometry: modular flow is Lorentz boost flow, and its KMS period becomes the
Unruh temperature along an accelerated orbit.

\begin{figure}[H]
  \centering
  \resizebox{0.80\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm,every node/.style={font=\scriptsize}]
    \tikzset{
      wedgebd/.style={line width=1.0pt},
      arrow/.style={-{Latex[length=2mm]},line width=0.85pt},
      box/.style={draw,rounded corners=4pt,line width=0.8pt,fill=orange!7,
        minimum width=2.8cm,minimum height=0.82cm,align=center}
    }
    \draw[->] (-0.85,0)--(3.35,0) node[right] {\(x^1\)};
    \draw[->] (0,-2.55)--(0,2.55) node[above] {\(x^0\)};
    \path[fill=blue!5] (0,0)--(2.62,2.62)--(2.62,-2.62)--cycle;
    \draw[wedgebd] (0,0)--(2.48,2.48)
      node[pos=0.86,above left,fill=white,inner sep=1pt] {\(x^0=x^1\)};
    \draw[wedgebd] (0,0)--(2.48,-2.48)
      node[pos=0.86,below left,fill=white,inner sep=1pt] {\(x^0=-x^1\)};
    \node[blue!55!black,fill=white,inner sep=1.5pt] at (2.12,0.04)
      {\(W_R=\{x^1>|x^0|\}\)};
    \draw[green!35!black,line width=0.6pt,domain=-1.05:1.05,samples=80,smooth,variable=\s]
      plot ({1.35*(exp(\s)+exp(-\s))/2},{1.35*(exp(\s)-exp(-\s))/2});
    \draw[arrow,green!45!black,domain=-0.95:0.95,samples=80,smooth,variable=\s]
      plot ({0.82*(exp(\s)+exp(-\s))/2},{0.82*(exp(\s)-exp(-\s))/2});
    \node[green!45!black,fill=white,inner sep=1.5pt] at (2.68,1.34)
      {boost orbit};
    \node[green!45!black,fill=white,inner sep=1.5pt] at (1.82,-1.36)
      {\((x^1)^2-(x^0)^2=\rho^2\)};
    \node[box] (modbox) at (-2.0,-1.55)
      {modular group\\\(\Delta_{W_R}^{\ii t}\)};
    \draw[arrow] (modbox.east) -- (0.73,-0.36);
    \node[fill=white,inner sep=1.5pt] at (-1.78,-2.35)
      {\(\Delta_{W_R}^{\ii t}=U(\Lambda_R(-2\pi t))\)};
  \end{tikzpicture}
  }
  \caption{By the Bisognano--Wichmann theorem, or by the corresponding explicit
  modular-covariance hypothesis in an AQFT net, wedge modular flow is geometric
  Lorentz boost flow; its flow lines inside the right wedge are hyperbolae
  \((x^1)^2-(x^0)^2=\rho^2\).}
  \label{fig:modular-wedge-flow}
\end{figure}

\section{Status of the Structural Inputs Used Above}
\label{sec:aqft-structural-input-status}

The preceding sections use several deep operator-algebraic and local-QFT
results.  Their logical statuses differ.  The distinction matters because an
abstract von Neumann algebra theorem may be used after standardness and
normality have been verified, whereas a local-QFT theorem usually requires
covariance, spectrum, locality, additivity, phase-space control, or
wedge-domain estimates in the concrete representation.

\begin{description}
\item[Doplicher--Roberts reconstruction.]
The purely categorical part says that a rigid symmetric \(C^*\)-tensor
category with simple unit, finite-dimensional morphism spaces, conjugates, and
finite statistical dimensions is equivalent to
\({\rm Rep}_{\rm fd}(G)\) for a compact group \(G\), after choosing a faithful
symmetric tensor fiber functor.  The local-QFT part is earlier: one must first
prove that the sectors are transportable bounded-region DHR endomorphisms,
that Haag duality places localized images and intertwiners in the assigned
local observable algebra rather than only in the dual net, and that locality
gives a symmetric exchange structure.  The local charged-field core
\eqref{eq:dr-local-field-core}--\eqref{eq:dr-local-haar-expectation} proves
the algebraic form of multiplication, fusion, conjugation, and fixed-point
projection once those hypotheses are present.  Charged fields for Gauss-law
sectors, spacelike-cone sectors, infraparticles, or confining theories require
different localization categories before a field-algebra theorem can even be
formulated.

\item[Tomita--Takesaki theory.]
This is pure operator-algebra infrastructure.  Its input is a standard pair
\((\mathcal R,\Omega)\), meaning \(\Omega\) is cyclic and separating for
\(\mathcal R\).  The chapter displays the \(S_0/F_0\) adjoint-core argument,
the closed Tomita operator, and the left-Hilbert-algebra mechanism leading to
\(\Delta^{\ii t}\mathcal R\Delta^{-\ii t}=\mathcal R\) and
\(J\mathcal R J=\mathcal R'\).  In QFT applications, the theorem may be used
only after standardness of the relevant local or wedge algebra has been
proved, typically through Reeh--Schlieder and locality.  It supplies an
abstract modular flow.  Identification of that flow with a spacetime symmetry
is an additional local-QFT theorem.

\item[Connes cocycle and Araki relative entropy.]
This is also operator-algebraic infrastructure, formulated for faithful normal
states or weights in a common standard representation.  The relative Tomita
operator and the product
\([D\psi:D\omega]_t=\Delta_{\psi|\omega}^{\ii t}\Delta_\omega^{-\ii t}\)
give an algebra unitary implementing the change of modular flow.  The finite
standard-form calculation fixes the density-matrix normalization.  In local
QFT the required work is to specify the von Neumann algebra, the normal states
on it, and the faithfulness/support condition.  Those data provide the
intrinsic type-III replacement for a formal density-matrix quotient.

\item[Nuclearity, split inclusions, and type classification.]
Buchholz--Wichmann nuclearity is a local phase-space hypothesis on
\(\Theta_{\beta,\mathcal O}:A\mapsto \ee^{-\beta H}A\Omega\).  The
operator-algebraic split criterion asks for normal extendability of the
multiplication representation
\(\mathcal M_1\odot\mathcal M_2'\to\mathcal B(\Hilb)\) to
\(\mathcal M_1\bar\otimes\mathcal M_2'\).  The text above explains how
summable separated normal functionals arise from energy damping, locality,
and the positive collar \(\mathcal O_1\Subset\mathcal O_2\).  The type
\(\mathrm{III}\) or type \(\mathrm{III}_1\) classification of sharp local
algebras is a separate theorem boundary: a type-I interpolant at positive
separation describes collar-separated independence, while type-III behavior
concerns the sharp zero-collar algebra.  Concrete models require their own
phase-space estimates, factoriality input, and modular-spectrum control.

\item[Borchers--Wiesbrock half-sided inclusions.]
The abstract theorem starts from a standard inclusion
\(\mathcal N\subset\mathcal M\) and a one-sided modular invariance condition.
Its content is the construction of a positive-generator translation group
with affine modular covariance.  The finite-dimensional rank argument above
isolates the content as infinite-dimensional standard-subspace geometry.  In
a Poincare-covariant QFT net, an inward translated wedge gives the geometric
half-sided inclusion only after wedge standardness, isotony, covariance, and
the relevant Bisognano--\allowbreak Wichmann modular identification are
available.  The light-ray
calculation then matches the abstract generator to the positive physical
momentum \(P^0-P^1\), while the converse reconstruction remains the
Borchers--Wiesbrock theorem itself.

\item[Bisognano--Wichmann and modular PCT.]
The Wightman polynomial-core theorem proved above establishes the boost KMS
strip from tube analyticity and locality.  The full wedge-modular theorem
additionally identifies the closure of the wedge Tomita operator with
geometric Lorentz boosts and the wedge reflection.  That
closed-operator/domain statement is the Bisognano--Wichmann QFT theorem
boundary unless the net is assumed to satisfy modular covariance.  Corollary
\ref{cor:aqft-pct-from-modular-covariance} is therefore conditional on the
explicit modular-covariance hypotheses stated there.
\end{description}

Thus each structural conclusion in this chapter is tied to a visible input:
DHR localization and duality for charge reconstruction; standardness for
modular theory; normal state data for relative entropy; phase-space bounds and
collars for split inclusions; half-sided standardness for
Borchers--Wiesbrock; and wedge-domain control for Bisognano--Wichmann.

\section{Algebraic Scattering Interface}

The Haag--Ruelle construction can be formulated in the algebraic setting using
almost-local operators whose energy-momentum transfer is concentrated near an
isolated mass shell.  Acting on the vacuum, such operators create
single-particle states.  Time-dependent translated operators with separated
velocity supports have limits that define incoming and outgoing scattering
states:
\[
  \ket{\psi_1,\ldots,\psi_n}_{\mathrm{in/out}}
  =
  \lim_{t\to\mp\infty}
  A_1(t)\cdots A_n(t)\vac .
\]
This gives the same conceptual order established earlier: first the
nonperturbative asymptotic states, then the scattering operator, and only
afterward any perturbative representation of matrix elements.

The algebraic language records the dependence of scattering on local
observables, spectral isolation, and phase-space properties.  Infraparticle
theories, gauge theories with long-range fields, and theories without a mass
gap require modified asymptotic constructions whose hypotheses must be stated
separately.
