\chapter{Global Forms and Higher-Form Symmetry}
\label{ch:global-forms-higher-form-symmetry}

\ConventionsMixed

The local Lie algebra of gauge transformations does not by itself define a
quantum gauge theory.  A continuum gauge theory also contains global bundle
data, a spectrum of genuine line and surface operators, and topological
defects that act on those extended operators.  This chapter gives the
finite-dimensional algebra that controls these data for compact connected
semisimple gauge groups.  The analytic definition of a renormalized
't Hooft line is given in
Chapter~\ref{ch:global-line-surface-domain-wall-operators}; here the purpose
is to define the electric and magnetic charge lattices on which that
construction depends.

\section{Character and Cocharacter Lattices}

Let \(\mathfrak g\) be a compact semisimple Lie algebra and let \(G_{\rm sc}\)
be the connected simply connected compact Lie group with Lie algebra
\(\mathfrak g\).  Fix a maximal torus \(T_{\rm sc}\subset G_{\rm sc}\), with
Lie algebra \(\mathfrak t\).  The character and cocharacter groups of a
compact torus \(T\) are
\[
  X^*(T)=\operatorname{Hom}_{\mathrm{Lie\,grp}}(T,U(1)),
  \qquad
  X_*(T)=\operatorname{Hom}_{\mathrm{Lie\,grp}}(U(1),T).
\]
They are free abelian groups of rank \(\dim T\).  A character
\(\lambda\in X^*(T)\) and a cocharacter \(m\in X_*(T)\) have an integer
pairing
\[
  \lambda\circ m:U(1)\longrightarrow U(1),
  \qquad
  z\longmapsto z^{\langle \lambda,m\rangle},
  \qquad
  \langle\lambda,m\rangle\in\mathbb Z .
\]
This pairing is part of the definition of \(X^*(T)\) and \(X_*(T)\); no
metric on \(\mathfrak t\) is needed.

For \(T_{\rm sc}\), write
\[
  \Lambda_{\rm wt}=X^*(T_{\rm sc}),\qquad
  \Lambda_{\rm crt}=X_*(T_{\rm sc}).
\]
These are the weight lattice and the coroot lattice.  The root lattice
\(\Lambda_{\rm rt}\subset\Lambda_{\rm wt}\) is generated by the roots of
\((G_{\rm sc},T_{\rm sc})\).  The coweight lattice is
\[
  \Lambda_{\rm cwt}=X_*(T_{\rm ad}),
  \qquad
  T_{\rm ad}=T_{\rm sc}/Z(G_{\rm sc}),
\]
and contains \(\Lambda_{\rm crt}\).  The finite quotients
\[
  \Lambda_{\rm wt}/\Lambda_{\rm rt},
  \qquad
  \Lambda_{\rm cwt}/\Lambda_{\rm crt}
\]
are naturally dual finite abelian groups.  The first quotient records
central characters of representations of \(G_{\rm sc}\); the second records
cocharacters modulo ordinary coroot magnetic charges.

\begin{definition}[Global form]
\label{def:compact-global-form}
A global form of the compact gauge group with Lie algebra \(\mathfrak g\) is
a connected compact group
\[
  G=G_{\rm sc}/\Gamma,
  \qquad
  \Gamma\subset Z(G_{\rm sc}).
\]
The maximal torus of \(G\) is \(T_G=T_{\rm sc}/\Gamma\).  Its character and
cocharacter lattices are
\begin{equation}
  X^*(T_G)
  =
  \{\lambda\in\Lambda_{\rm wt}\mid \lambda(\gamma)=1
  \text{ for every }\gamma\in\Gamma\},
  \label{eq:global-form-character-lattice}
\end{equation}
and
\begin{equation}
  X_*(T_G)
  =
  \{m\in\Lambda_{\rm cwt}\mid
  \exp_{G_{\rm sc}}(2\pi\ii m)\in\Gamma\}.
  \label{eq:global-form-cocharacter-lattice}
\end{equation}
The exponential in the second formula uses the identification of a
cocharacter with an element of \(\mathfrak t\) whose exponential closes in
the adjoint torus.  It closes in \(G=G_{\rm sc}/\Gamma\) precisely when its
endpoint in \(G_{\rm sc}\) lies in \(\Gamma\).
\end{definition}

The local curvature term in the Yang--Mills Lagrangian depends only on
\(\mathfrak g\) and on the normalization of the invariant quadratic form.
The global form changes the allowed bundles and the admissible line
operators.  These changes are invisible in ordinary perturbation theory
around a trivial bundle unless extended operators or background fields for
higher-form symmetries are included.

Let \(V_\lambda\) be an irreducible representation of \(G_{\rm sc}\) with
highest weight \(\lambda\in\Lambda_{\rm wt}\).  It defines a representation
of \(G=G_{\rm sc}/\Gamma\) if and only if
\(\lambda\in X^*(T_G)\), equivalently if and only if every element of
\(\Gamma\) acts trivially on \(V_\lambda\).
The representation \(V_\lambda\) factors through the quotient map
\(G_{\rm sc}\to G_{\rm sc}/\Gamma\) exactly when \(\Gamma\) is contained in
the kernel of the representation.  Since \(\Gamma\) is central, Schur's
lemma says that every \(\gamma\in\Gamma\) acts on the irreducible module
\(V_\lambda\) by a scalar.  That scalar is the central character determined
by \(\lambda\), namely the restriction of \(\lambda\) to
\(Z(G_{\rm sc})\subset T_{\rm sc}\).  Hence the action of every
\(\gamma\in\Gamma\) is trivial exactly when
\(\lambda(\gamma)=1\) for all \(\gamma\), which is
\eqref{eq:global-form-character-lattice}.

\begin{example}[\(SU(N)/\mathbb Z_k\)]
\label{ex:su-n-z-k-descent}
Let \(k\) divide \(N\), let \(z=\exp(2\pi\ii/N)\mathbf 1_N\), and let
\(\mathbb Z_k\subset Z(SU(N))\) be the subgroup generated by \(z^{N/k}\).
An irreducible \(SU(N)\)-representation has an \(N\)-ality
\[
  e\in\mathbb Z_N=\mathbb Z/N\mathbb Z,
\]
defined as the number of boxes of its Young diagram modulo \(N\), or
equivalently as its central character
\(z\mapsto\exp(2\pi\ii e/N)\).  The representation descends to
\(SU(N)/\mathbb Z_k\) if and only if
\[
  \exp(2\pi\ii e/k)=1,
  \qquad\text{that is}\qquad
  e\in k\mathbb Z_N .
\]
Thus the quotient removes Wilson lines whose \(N\)-ality is not divisible by
\(k\).  Dually, the cocharacter lattice is enlarged: modulo the coroot
lattice, the allowed magnetic classes are
\[
  m\in (N/k)\mathbb Z_N .
\]
Indeed, after choosing the first fundamental coweight as the generator of
\(\Lambda_{\rm cwt}/\Lambda_{\rm crt}\cong\mathbb Z_N\), the cocharacter
represented by \(m\) times this generator closes in \(SU(N)/\mathbb Z_k\)
precisely when its endpoint \(z^m\) lies in \(\mathbb Z_k\), i.e. precisely
when \(m\) is a multiple of \(N/k\) modulo \(N\).
\end{example}

\section{Genuine Lines and the Dirac Pairing}

\begin{definition}[Genuine line operator]
\label{def:genuine-line-operator}
A line operator \(L(C)\) supported on an embedded oriented curve \(C\) is
genuine if its definition does not require a choice of an auxiliary surface
ending on \(C\).  If a surface \(S\) with \(\partial S=C\) is part of the
definition, then the insertion is a boundary of a higher-dimensional defect
rather than an autonomous line operator.  In gauge theory, being genuine is a
property of the chosen global form, discrete theta data, and spectrum of
allowed charged defects; it is not a property of the local gauge algebra
alone.
\end{definition}

Before dividing by screening, a dyonic line in a gauge theory with local
algebra \(\mathfrak g\) has an electric weight and a magnetic cocharacter.
For the finite center-sensitive part of the discussion one may record only
the classes
\[
  \overline\lambda\in\Lambda_{\rm wt}/\Lambda_{\rm rt},
  \qquad
  \overline m\in\Lambda_{\rm cwt}/\Lambda_{\rm crt}.
\]
The root and coroot quotients remove charges that can be screened by adjoint
gluonic fields or shifted by ordinary smooth monopole cores.  The remaining
finite data are precisely the charges detected by center and magnetic
one-form symmetries.

For two dyonic labels
\[
  \gamma=(\lambda,m),
  \qquad
  \gamma'=(\lambda',m'),
\]
define the Dirac pairing
\begin{equation}
  B(\gamma,\gamma')
  =
  \lambda(m')-\lambda'(m)
  \quad\in\mathbb Q .
  \label{eq:dirac-pairing-lattice}
\end{equation}
The value may be fractional when \(\lambda\) is a weight and \(m'\) is a
coweight of the universal center-sensitive lattice.  Its class in
\(\mathbb Q/\mathbb Z\) is independent of representatives modulo
\(\Lambda_{\rm rt}\oplus\Lambda_{\rm crt}\).  Two lines can be simultaneously
genuine in the same theory only if
\begin{equation}
  B(\gamma,\gamma')\in\mathbb Z .
  \label{eq:mutual-locality-dirac-condition}
\end{equation}
Equivalently, the finite Aharonov--Bohm phase
\(\exp(2\pi\ii B(\gamma,\gamma'))\) must be trivial.

\begin{definition}[Finite line-spectrum datum]
\label{def:finite-line-spectrum-datum}
For a fixed local gauge algebra, the finite line-spectrum datum is a subgroup
\[
  L\subset
  \bigl(\Lambda_{\rm wt}/\Lambda_{\rm rt}\bigr)
  \oplus
  \bigl(\Lambda_{\rm cwt}/\Lambda_{\rm crt}\bigr)
\]
on which the Dirac pairing \eqref{eq:dirac-pairing-lattice} vanishes modulo
\(\mathbb Z\).  It is maximal if it equals its orthogonal complement with
respect to that finite pairing.  Maximality is the finite-charge statement
that no further center-sensitive Wilson--'t Hooft charge can be added while
preserving mutual locality.
\end{definition}

The definition is deliberately finite and algebraic.  A complete continuum
definition of the corresponding line operator also requires singular
boundary conditions, renormalization of the defect, and a regulator or
constructive framework; these analytic data are separated in
Definition~\ref{def:renormalized-thooft-line}.

\begin{proposition}[\(\mathfrak{su}(N)\) finite line lattices]
\label{prop:su-n-finite-line-lattices}
For local algebra \(\mathfrak{su}(N)\), identify
\[
  \Lambda_{\rm wt}/\Lambda_{\rm rt}
  \cong
  \Lambda_{\rm cwt}/\Lambda_{\rm crt}
  \cong
  \mathbb Z_N.
\]
Write a finite dyonic charge as \((e,m)\in\mathbb Z_N\oplus\mathbb Z_N\).
The finite Dirac pairing is
\begin{equation}
  \overline B\bigl((e,m),(e',m')\bigr)
  =
  \frac{em'-e'm}{N}
  \quad\in\mathbb Q/\mathbb Z .
  \label{eq:su-n-finite-dirac-pairing}
\end{equation}
Let \(k\mid N\) and let \(p\in\mathbb Z_k\).  The subgroup
\begin{equation}
  L_{N,k,p}
  =
  \bigl\langle (k,0),(p,N/k)\bigr\rangle
  \subset \mathbb Z_N\oplus\mathbb Z_N
  \label{eq:su-n-z-k-line-lattice}
\end{equation}
is maximal isotropic.  In the Yang--Mills convention used in
Chapter~\ref{ch:global-discrete-theta-terms}, it is the finite
Wilson--'t Hooft line lattice of the \(SU(N)/\mathbb Z_k\) theory with
discrete theta parameter \(p\), before adding matter that could screen
center charges.
\end{proposition}

\begin{proof}
The quotient
\(\Lambda_{\rm wt}/\Lambda_{\rm rt}\cong\mathbb Z_N\) is generated by the
fundamental Wilson \(N\)-ality.  The dual quotient
\(\Lambda_{\rm cwt}/\Lambda_{\rm crt}\cong\mathbb Z_N\) is generated by the
fundamental coweight magnetic class.  The generator pair has pairing
\(1/N\), which gives \eqref{eq:su-n-finite-dirac-pairing} by bilinearity and
antisymmetry.

The two generators in \eqref{eq:su-n-z-k-line-lattice} pair as
\[
  \overline B((k,0),(p,N/k))=1=0\quad\text{in }\mathbb Q/\mathbb Z,
\]
and each generator pairs trivially with itself by antisymmetry.  Hence
\(L_{N,k,p}\) is isotropic.  Its first generator gives \(N/k\) distinct
electric classes.  Modulo this subgroup, the second generator has magnetic
component \(N/k\), which has order \(k\).  Thus \(|L_{N,k,p}|=N\).

It remains to see that no larger isotropic subgroup exists.  If
\((x,y)\in(\mathbb Z_N)^2\) pairs trivially with \((k,0)\), then
\(-ky/N\in\mathbb Z\), so \(y=(N/k)b\) modulo \(N\).  Trivial pairing with
\((p,N/k)\) then gives
\[
  \frac{x(N/k)-py}{N}
  =
  \frac{x-pb}{k}
  \in\mathbb Z,
\]
so \(x=pb+ka\) modulo \(N\).  Therefore
\[
  (x,y)=a(k,0)+b(p,N/k),
\]
and the orthogonal complement of \(L_{N,k,p}\) is \(L_{N,k,p}\) itself.
This proves maximality.  The identification with the global form follows
from Example~\ref{ex:su-n-z-k-descent}: Wilson charges are multiples of
\(k\), magnetic charges are multiples of \(N/k\), and the discrete theta
coordinate shifts the minimal magnetic line by electric charge \(p\).
\end{proof}

Figure~\ref{fig:su5-line-charge-lattice} displays the finite line-charge
lattice in the \(SU(5)\)-type example and marks the maximal mutually local
sublattice selected by the global form and discrete theta coordinate.

\begin{figure}[H]
\centering
\begin{tikzpicture}[scale=0.78, every node/.style={font=\small}]
  \draw[->] (-0.35,0) -- (5.6,0) node[right] {\(e\)};
  \draw[->] (0,-0.35) -- (0,5.1) node[above] {\(m\)};
  \foreach \x in {0,...,4} {
    \foreach \y in {0,...,4} {
      \fill[black!35] (\x,\y) circle (1.3pt);
    }
  }
  \foreach \x in {0,...,4} {
    \fill[blue!70!black] (\x,0) circle (2.4pt);
  }
  \foreach \y in {0,...,4} {
    \fill[red!75!black] (0,\y) circle (2.4pt);
  }
  \node[blue!70!black,below] at (2,-0.18) {\(SU(5):\ L=\langle(1,0)\rangle\)};
  \node[red!75!black,rotate=90,above] at (-0.18,2.45)
    {\(PSU(5):\ L=\langle(0,1)\rangle\)};
  \draw[rounded corners,blue!70!black] (-0.12,-0.12) rectangle (4.12,0.12);
  \draw[rounded corners,red!75!black] (-0.12,-0.12) rectangle (0.12,4.12);
\end{tikzpicture}
\caption{The finite \(\mathfrak{su}(5)\) charge group
\(\mathbb Z_5^{\mathrm e}\oplus\mathbb Z_5^{\mathrm m}\).  A line spectrum is not the
whole grid but a maximal isotropic subgroup for
\eqref{eq:su-n-finite-dirac-pairing}.  The \(SU(5)\) and \(PSU(5)\) choices
shown here are the two coordinate axes; discrete theta choices tilt the
magnetic axis as in \eqref{eq:su-n-z-k-line-lattice}.}
\label{fig:su5-line-charge-lattice}
\end{figure}

\section{Abelian Group-Like Higher-Form Symmetry}

\begin{definition}[Abelian group-like \(p\)-form global symmetry]
\label{def:p-form-global-symmetry}
Let \(D\) be the spacetime dimension and let \(0\leq p\leq D-2\).  A
group-like abelian \(p\)-form global symmetry with finite or compact abelian
group \(A\) assigns to every closed oriented codimension-\((p+1)\)
submanifold \(M^{D-p-1}\) and every \(a\in A\) an invertible topological defect
\[
  U_a(M).
\]
The defects obey
\[
  U_a(M)U_b(M)=U_{a+b}(M),
  \qquad
  U_0(M)=1,
\]
inside correlation functions, and their correlation functions are invariant
under smooth deformations of \(M\) that avoid charged insertions, boundaries,
and singularities.  A charged \(p\)-dimensional operator
\(W_\chi(\Sigma^p)\) has charge
\(\chi\in A^\vee=\operatorname{Hom}_{\mathrm{cont}}(A,U(1))\) if
\begin{equation}
  U_a(M)W_\chi(\Sigma)
  =
  \chi(a)^{\operatorname{Link}(M,\Sigma)}
  W_\chi(\Sigma)
  \label{eq:higher-form-linking-action}
\end{equation}
inside correlation functions whenever \(M\) and \(\Sigma\) are disjoint and
their linking number is defined.
\end{definition}

For \(p=0\), this definition covers the abelian part of ordinary internal
symmetry.  Nonabelian zero-form symmetry requires topological defects labelled
by group elements with noncommutative fusion; the one-form symmetries used
below are abelian in precisely the sense of
Definition~\ref{def:p-form-global-symmetry}.

For a continuous \(U(1)\) \(p\)-form symmetry, a charge is an integer
\(q\in\mathbb Z\), the character is \(\chi_q(e^{\ii\alpha})=e^{\ii q\alpha}\),
and \eqref{eq:higher-form-linking-action} becomes
\[
  U_\alpha(M)W_q(\Sigma)
  =
  \exp\!\bigl(
    \ii\alpha q\,\operatorname{Link}(M,\Sigma)
  \bigr)W_q(\Sigma).
\]
The linking formula is not an analogy: it is the operator definition of the
symmetry action on extended charged objects.

\section{Electric-Magnetic One-Form Symmetries}

Let \(G\) be a compact connected gauge group and assume for the moment that
there is no dynamical matter carrying nontrivial center charge and no
dynamical magnetic object that allows 't Hooft lines to end.  The group of
electric one-form symmetry defects is
\[
  Z(G).
\]
Its characters act on Wilson lines.  A Wilson line in representation \(R\)
has electric one-form charge given by the central character
\[
  Z(G)\longrightarrow U(1),
  \qquad
  z\longmapsto R(z).
\]
If \(G=SU(N)\), this is the usual \(N\)-ality.  If matter of nonzero
\(N\)-ality is present, the corresponding Wilson line can end on a dynamical
field and the electric one-form symmetry is explicitly reduced or broken.

For a non-simply connected \(G\), there is also a magnetic one-form symmetry
whose background fields measure the obstruction class of \(G\)-bundles.
The topological magnetic surface defects are labelled by
\[
  \pi_1(G)^\vee=\operatorname{Hom}(\pi_1(G),U(1)).
\]
Equivalently, a background two-form field for this symmetry is valued in
\(\pi_1(G)\).  The distinction between \(\pi_1(G)\) and its Pontryagin dual
is only invisible after choosing a noncanonical perfect pairing; the
definition keeps them separate.  For \(G=SU(N)/\mathbb Z_k\),
\[
  Z(G)\cong \mathbb Z_{N/k},
  \qquad
  \pi_1(G)\cong \mathbb Z_k .
\]
The line lattice \(L_{N,k,p}\) in
Proposition~\ref{prop:su-n-finite-line-lattices} is the finite algebraic
record of which Wilson--'t Hooft probes are genuine for the chosen global
form and discrete theta parameter.

Screening and confinement are therefore statements about the behavior of
renormalized extended operators with charges in a specified line spectrum.
They are developed in
Chapter~\ref{ch:global-confinement-screening-oblique}.  Gauging a finite
one-form symmetry and tracking its anomaly require cochain-level background
fields and are developed in
Chapters~\ref{ch:global-discrete-theta-terms} and
\ref{ch:global-higher-group-symmetry-tft}.

\begin{openproblem}[Global structure from continuum local data]
\label{op:global-structure-from-local-data}
Determine, for a continuum QFT specified by local fields and local
correlators, what additional data recover the full lattice of genuine line,
surface, and defect operators.  A Wightman field presentation determines
local correlators and charged local sectors under stated hypotheses; global
form, higher-form symmetries, and extended-operator completeness require
further information about bundles, boundary conditions around defects,
screening, and topological defect fusion.
\end{openproblem}
