\chapter{Discrete Theta Terms}
\label{ch:global-discrete-theta-terms}

\ConventionsEuclidean

Discrete theta terms are topological phases depending on global bundle data
and finite higher-form gauge fields.  They are part of the definition of a
gauge theory with a specified global form: the same local gauge algebra and
the same curvature Lagrangian can have distinct sums over bundles, distinct
topological weights, and distinct spectra of genuine line operators.  This
chapter gives the cochain-level definitions needed for the four-dimensional
gauge-theory applications in Volume IX.

\section{Finite Cochains And Topological Actions}

Let \(M\) be a closed oriented triangulated \(D\)-manifold.  For an abelian
group \(A\), write
\[
  C^r(M,A),\qquad Z^r(M,A),\qquad B^r(M,A),\qquad H^r(M,A)
\]
for \(r\)-cochains, cocycles, coboundaries, and cohomology classes.  The
coboundary is denoted \(\delta\).  A finite \(A\)-valued zero-form gauge
field is a one-cocycle
\[
  a\in Z^1(M,A),
\]
with gauge equivalence
\[
  a\sim a+\delta\lambda,\qquad \lambda\in C^0(M,A).
\]
For a finite \(p\)-form symmetry or gauge field, the background is instead a
\((p+1)\)-cocycle
\[
  B\in Z^{p+1}(M,A),
  \qquad
  B\sim B+\delta\Lambda,\quad \Lambda\in C^p(M,A).
\]

An exponentiated finite topological action is a function
\[
  S_{\mathrm{top}}:Z^{p+1}(M,A)\longrightarrow U(1)
\]
that is invariant under subdivision and gauge transformation, and is local in
the sense that it is obtained by evaluating a \(D\)-cocycle built from
cochains on the fundamental class of \(M\).  For a zero-form finite gauge
field this data is equivalently represented by a class
\[
  [\omega]\in H^D(BA,U(1)).
\]
For a \(p\)-form field it is represented by a class on \(B^{p+1}A\).
Explicitly, if \(\omega(a)\in C^D(M,\mathbb R/\mathbb Z)\) is a cocycle
depending naturally on \(a\), then
\[
  S_{\mathrm{top}}[a]
  =
  \exp\left(2\pi\ii\int_M\omega(a)\right).
\]
Here
\[
  \int_M:C^D(M,\mathbb R/\mathbb Z)\longrightarrow\mathbb R/\mathbb Z
\]
is evaluation on the fundamental cycle.

\begin{proposition}[Coboundary invariance on closed manifolds]
\label{prop:discrete-theta-coboundary-invariance}
If two finite topological Lagrangians differ by a coboundary,
\[
  \omega'(a)-\omega(a)=\delta\alpha(a),
\]
then their exponentiated actions agree on every closed \(D\)-manifold for
every flat finite background field \(a\).
\end{proposition}

\begin{proof}
Let \(c_M\) be a fundamental \(D\)-cycle of \(M\).  Since \(M\) is closed,
\[
  \partial c_M=0.
\]
The evaluation pairing satisfies
\[
  \langle\delta\alpha(a),c_M\rangle
  =
  \langle\alpha(a),\partial c_M\rangle
  =
  0
  \quad\text{in }\mathbb R/\mathbb Z .
\]
Therefore
\[
  \exp\left(2\pi\ii\int_M\omega'(a)\right)
  =
  \exp\left(2\pi\ii\int_M\omega(a)\right).
\]
For a manifold with boundary the same calculation leaves the boundary
evaluation \(\int_{\partial M}\alpha(a)\), which is the boundary datum used in
anomaly inflow.
\end{proof}

\section{Two-Form \texorpdfstring{\(\mathbb Z_N\)}{ZN} Fields}

The four-dimensional examples use a two-form finite field
\[
  B\in Z^2(M,\mathbb Z_N).
\]
Choose an integral lift
\[
  \widetilde B\in C^2(M,\mathbb Z)
\]
whose reduction modulo \(N\) is \(B\).  Since \(B\) is closed modulo \(N\),
there is an integral three-cochain \(C\) such that
\[
  \delta\widetilde B=N C .
\]
The Pontryagin square is the cohomology operation
\[
  \mathcal P:H^2(M,\mathbb Z_N)\longrightarrow H^4(M,\mathbb Z_{2N})
\]
characterized by
\[
  \mathcal P(B)\equiv B\smile B \pmod N
\]
and by the quadratic-refinement identity
\[
  \mathcal P(B+B')
  =
  \mathcal P(B)+\mathcal P(B')+2\,B\smile B'
  \quad\text{in }H^4(M,\mathbb Z_{2N}).
\]
A cochain representative is
\[
  \mathcal P(B)
  =
  \widetilde B\smile\widetilde B
  -N\,\widetilde B\smile_1 C
  \quad\text{mod }2N,
\]
where \(\smile_1\) is Steenrod's first cup product.  The second term is what
makes the expression independent of the chosen lift and invariant under
\(B\mapsto B+\delta\Lambda\).  When \(N\) is odd, \(2\) is invertible modulo
\(N\), and the same information can be represented by the ordinary square
\[
  B\smile B\in H^4(M,\mathbb Z_N).
\]
When \(N\) is even, the lift to \(\mathbb Z_{2N}\) is essential.

The corresponding four-dimensional local counterterms for a background
\(\mathbb Z_N\) one-form symmetry have the form
\[
  S_p[B]
  =
  \exp\left(
    \frac{2\pi\ii p}{2N}
    \int_M\mathcal P(B)
  \right).
\]
On oriented manifolds the coefficient is valued in
\[
  p\in
  \begin{cases}
    \mathbb Z_N, & N\ \text{odd},\\
    \mathbb Z_{2N}, & N\ \text{even},
  \end{cases}
\]
using the identifications of \(H^4(B^2\mathbb Z_N,U(1))\).  On spin
manifolds some terms that differ on a general oriented manifold can coincide,
because the Wu-class relations constrain the integral of the Pontryagin
square.  For odd \(N\), the displayed formula is read as
\[
  \frac{p}{2N}\mathcal P(B)
  =
  \frac{p}{N}\,\frac{1}{2}B\smile B
  \quad\text{in }\mathbb R/\mathbb Z,
\]
where \(1/2\) denotes the inverse of \(2\) in \(\mathbb Z_N\).  Thus the
spacetime category is part of the definition of the available counterterm
group.

\begin{proposition}[Quadratic refinement phase]
\label{prop:pontryagin-square-quadratic-phase}
Let \(B,B'\in H^2(M,\mathbb Z_N)\).  The ratio of counterterms obeys
\[
  \frac{S_p[B+B']}{S_p[B]S_p[B']}
  =
  \exp\left(
    \frac{2\pi\ii p}{N}
    \int_M B\smile B'
  \right).
\]
\end{proposition}

\begin{proof}
Use the defining identity of the Pontryagin square:
\[
  \mathcal P(B+B')-\mathcal P(B)-\mathcal P(B')
  =
  2B\smile B'
  \quad\text{mod }2N .
\]
Substituting this into the exponent gives
\[
  \frac{2\pi\ii p}{2N}
  \int_M 2B\smile B'
  =
  \frac{2\pi\ii p}{N}
  \int_M B\smile B' .
\]
Exponentiation gives the stated ratio.
\end{proof}

\section{Global Form And Obstruction Classes}

Let
\[
  G=SU(N)/\mathbb Z_k,\qquad k\mid N,
\]
where \(\mathbb Z_k\subset Z(SU(N))\) is generated by
\[
  \exp(2\pi\ii/k)\mathbf 1_N=\exp(2\pi\ii\,N/kN)\mathbf 1_N .
\]
A principal \(G\)-bundle \(P_G\to M\) has an obstruction class
\[
  w_2^{(k)}(P_G)\in H^2(M,\mathbb Z_k)
\]
to lifting \(P_G\) to an \(SU(N)\)-bundle.  This class is obtained from the
central extension
\[
  1\longrightarrow\mathbb Z_k
  \longrightarrow SU(N)
  \longrightarrow SU(N)/\mathbb Z_k
  \longrightarrow1
\]
by the connecting homomorphism in nonabelian cohomology.  Concretely, choose
local lifts of the \(G\)-transition functions to \(SU(N)\).  On triple
overlaps the lifted transition functions multiply to a \(\mathbb Z_k\)-valued
two-cocycle.  Its cohomology class is \(w_2^{(k)}(P_G)\).  Changing the local
lifts changes this two-cocycle by a coboundary.

For \(k=N\), \(G=PSU(N)\) and the obstruction class is usually written
\[
  w_2(P_{PSU(N)})\in H^2(M,\mathbb Z_N).
\]

\begin{proposition}[Fractional \(PSU(N)\) instanton number]
\label{prop:psu-n-fractional-instanton-number}
With the instanton-number convention compatible with
\[
  k=c_2(E)
\]
for an \(SU(N)\)-bundle \(E\), the fractional part of the instanton number of
a \(PSU(N)\)-bundle is fixed by its obstruction class:
\[
  \nu(P_{PSU(N)})
  =
  -\frac{N-1}{2N}\int_M\mathcal P(w_2)
  \quad\text{mod }1 .
\]
\end{proposition}

\begin{proof}
A projective \(PSU(N)\)-bundle can be described locally by \(SU(N)\)
transition functions that close on triple overlaps up to a central
\(\mathbb Z_N\)-valued two-cocycle representing \(w_2\).  Equivalently, for
the purpose of computing the fractional characteristic class, choose local
\(U(N)\) lifts and write \(x\) for the integral Cech two-cocycle that reduces
to \(w_2\) modulo \(N\).  If the projective bundle is represented by an
ordinary \(U(N)\)-bundle \(E\), then \(x=c_1(E)\); the local Cech computation
gives the same expression in general.

The projective second Chern coordinate is
\[
  \nu(E)
  =
  c_2(E)-\frac{N-1}{2N}c_1(E)^2
  \quad\in\mathbb Q .
\]
This combination is invariant under tensoring \(E\) by a line bundle.  Indeed,
if \(\ell=c_1(L)\), then
\[
  c_1(E\otimes L)=c_1(E)+N\ell
\]
and
\[
  c_2(E\otimes L)
  =
  c_2(E)+(N-1)c_1(E)\ell+\frac{N(N-1)}{2}\ell^2 .
\]
Substitution gives
\[
\begin{aligned}
  c_2(E\otimes L)
  &-\frac{N-1}{2N}c_1(E\otimes L)^2\\
  &=
  c_2(E)-\frac{N-1}{2N}c_1(E)^2 .
\end{aligned}
\]
Thus \(\nu\) is a characteristic class of the projective bundle.  Since
\(c_2(E)\) is integral, its fractional part is
\[
  -\frac{N-1}{2N}c_1(E)^2
  \quad\text{mod }1 .
\]
The Pontryagin square is the lift of \(w_2^2\) modulo \(2N\); hence
\[
  c_1(E)^2\equiv \mathcal P(w_2)\pmod{2N}.
\]
The displayed formula follows.
\end{proof}

This characteristic-class formula is the global reason that shifting the
ordinary continuous theta angle by \(2\pi\) changes the discrete theta
coordinate in a theory whose bundles fail to lift to \(SU(N)\).

\section{Discrete Theta Coordinate For \texorpdfstring{\(SU(N)/\mathbb Z_k\)}{SU(N)/Zk}}

The pure \(SU(N)/\mathbb Z_k\) Yang--Mills theories have a discrete theta
coordinate
\[
  p\in\mathbb Z_k
\]
in addition to the continuous theta angle.  This coordinate can be detected
without writing a continuum path integral: it is the electric tilt of the
minimal magnetic line.  With the finite charge notation of
Chapter~\ref{ch:global-forms-higher-form-symmetry}, write a dyonic line label
as
\[
  (e,m)\in\mathbb Z_N^{\mathrm e}\oplus\mathbb Z_N^{\mathrm m}.
\]
The genuine line lattice is
\[
  L_{N,k,p}
  =
  \langle (k,0),(p,N/k)\rangle .
\]
The generator \((k,0)\) records the Wilson lines that descend to
\(SU(N)/\mathbb Z_k\).  The generator \((p,N/k)\) records the minimal magnetic
line, whose magnetic charge is \(N/k\) and whose electric charge is shifted by
the discrete theta coordinate \(p\).  The identification
\[
  p\sim p+k
\]
holds because
\[
  (p+k,N/k)=(p,N/k)+(k,0).
\]
Thus \(p\) is naturally a class in \(\mathbb Z_k\).

\begin{proposition}[Discrete theta tilt and maximal locality]
\label{prop:discrete-theta-line-tilt}
For each \(p\in\mathbb Z_k\), the line lattice
\[
  L_{N,k,p}
  =
  \langle (k,0),(p,N/k)\rangle
  \subset\mathbb Z_N\oplus\mathbb Z_N
\]
is maximal isotropic for the finite Dirac pairing
\[
  \overline B((e,m),(e',m'))
  =
  \frac{em'-e'm}{N}
  \quad\in\mathbb Q/\mathbb Z .
\]
Changing \(p\) changes the electric charge of the minimal magnetic line by a
Witten-effect shift.
\end{proposition}

\begin{proof}
The isotropy and maximality are proved in
Proposition~\ref{prop:su-n-finite-line-lattices}.  The new point is the
interpretation of \(p\).  The minimal magnetic generator has magnetic
component \(N/k\), the generator of the magnetic subgroup allowed by the
global form.  Its electric component is \(p\).  Replacing \(p\) by \(p+k\)
adds the Wilson generator \((k,0)\), so the subgroup is unchanged.  Therefore
the inequivalent tilts are indexed by \(p\in\mathbb Z_k\).  A shift of the
continuous theta angle by \(2\pi\) changes the electric charge of a magnetic
line by its magnetic charge in the finite pairing convention; in these
coordinates this sends
\[
  p\longmapsto p+1
  \quad\text{in }\mathbb Z_k .
\]
\end{proof}

For \(N=2\) this gives the familiar three finite line spectra:
\[
  SU(2):\ \langle(1,0)\rangle,\qquad
  SO(3)_+:\ \langle(0,1)\rangle,\qquad
  SO(3)_-:\ \langle(1,1)\rangle .
\]
The last two have the same local Lie algebra and the same quotient gauge
group \(SO(3)=SU(2)/\mathbb Z_2\), but different discrete theta coordinate
and different dyonic line spectrum.

\section{Gauging One-Form Symmetry And Counterterms}

Start with the \(SU(N)\) theory and its electric
\(\mathbb Z_N\) one-form symmetry.  Coupling this symmetry to a background
two-form field
\[
  B\in Z^2(M,\mathbb Z_N)
\]
modifies the allowed bundles by allowing transition functions to close only
up to the center.  Gauging a subgroup \(\mathbb Z_k\subset\mathbb Z_N\) means
summing over the corresponding two-form gauge field, with a choice of
Pontryagin-square counterterm.  The result is an
\(SU(N)/\mathbb Z_k\) theory, and the counterterm choice is the discrete theta
coordinate \(p\) seen by the line lattice above.

The counterterm also controls anomalies.  If the generating functional in
background \(B\) transforms under a one-form gauge transformation by a phase
\[
  Z[B+\delta\Lambda]
  =
  \exp\left(2\pi\ii\int_M \alpha(B,\Lambda)\right)Z[B],
\]
then the phase is removed by a local counterterm exactly when
\(\alpha\) is the gauge variation of a \(D\)-dimensional local cochain.  If
the phase is the boundary variation of a \((D+1)\)-dimensional invertible
theory, it is an anomaly.  Discrete theta terms and anomaly inflow are
therefore the same cohomological mechanism viewed in neighboring dimensions:
in \(D\) dimensions the cochain changes the definition of the theory, and in
\(D+1\) dimensions its variation supplies an inflow phase for a boundary.

\section{Continuum Gauge-Theory Status}

A continuum Yang--Mills theory with nontrivial global form is specified by:
\[
\begin{array}{ll}
  \text{local data} &
  \mathfrak{su}(N)\text{ fields and curvature action},\\[1mm]
  \text{global form} &
  SU(N)/\mathbb Z_k,\\[1mm]
  \text{bundle sum} &
  \text{principal }SU(N)/\mathbb Z_k\text{ bundles on }M,\\[1mm]
  \text{topological weights} &
  \theta\text{ and }p\in\mathbb Z_k,\\[1mm]
  \text{line spectrum} &
  L_{N,k,p}\subset\mathbb Z_N^{\mathrm e}\oplus\mathbb Z_N^{\mathrm m},\\[1mm]
  \text{background fields} &
  \text{two-form backgrounds for remaining one-form symmetries}.
\end{array}
\]
A lattice regulator implements the same data by plaquette variables and
finite cochains for higher-form backgrounds.  Matching a lattice construction
to the continuum description requires matching the obstruction class
\(w_2^{(k)}\), the Pontryagin-square phase, the continuous theta coordinate,
and the genuine line spectrum.  Perturbation theory around the trivial bundle
sees only the local Lie algebra and curvature vertices; the discrete theta
data are detected by global bundle sectors and extended operators.

\begin{openproblem}[Discrete theta terms in interacting continuum QFT]
\label{op:discrete-theta-continuum-qft}
Give a nonperturbative construction of four-dimensional gauge theories with
all global forms and discrete theta terms in which bundles, one-form
backgrounds, line operators, anomalies, and continuum limits are treated in a
single framework, and prove the equivalence between continuum characteristic
classes, lattice finite-cochain definitions, and the operator line spectra.
\end{openproblem}
