\chapter{Anomaly Inflow and Invertible Field Theories}
\label{ch:global-anomaly-inflow-invertible-field-theories}

\ConventionsMixed

An anomaly is a precise failure of background-field covariance.  The
partition function of an anomalous \(D\)-dimensional theory is not an
ordinary function on the moduli groupoid of background fields; it is a vector
in a one-dimensional complex vector space depending on the background.  An
invertible \((D+1)\)-dimensional theory packages these lines and their
gluing laws.  Inflow is the statement that the anomalous \(D\)-dimensional
theory is a boundary condition, or relative theory, for this invertible bulk.

\begin{definition}[Anomaly-inflow structure]
\label{def:anomaly-inflow-datum}
Fix a spacetime dimension \(D\).  An anomaly-inflow structure is a tuple
\[
  \mathfrak A
  =
  \bigl(\mathcal B,\mathrm{Bord}_{D+1}^{\mathcal B},
  \mathcal Z_{\mathrm{bulk}},\mathcal L,Z_{\partial}\bigr)
\]
with the following entries.
\begin{enumerate}
\item \(\mathcal B\) assigns to each closed \(D\)-manifold \(M\), with the
declared spacetime structure, a groupoid \(\mathcal B(M)\) of background
fields and background isomorphisms.
\item \(\mathrm{Bord}_{D+1}^{\mathcal B}\) is the corresponding background
bordism category.  Its objects are pairs \((M,A)\), \(A\in\mathcal B(M)\);
its morphisms are \((D+1)\)-dimensional bordisms \(X:M_0\to M_1\) equipped
with background fields restricting to the prescribed boundary backgrounds.
\item \(\mathcal Z_{\mathrm{bulk}}\) is an invertible \((D+1)\)-dimensional
field theory,
\[
  \mathcal Z_{\mathrm{bulk}}:
  \mathrm{Bord}_{D+1}^{\mathcal B}\longrightarrow
  \mathrm{Lines}_{\mathbb C},
\]
namely a symmetric monoidal functor to one-dimensional complex vector spaces
and nonzero linear maps.  On a closed \((D+1)\)-manifold \(X\) it gives a
nonzero complex number.
\item \(\mathcal L\) is the boundary line functor induced by
\(\mathcal Z_{\mathrm{bulk}}\).  For each closed \(M\) it is a functor
\[
  \mathcal L_M:\mathcal B(M)\longrightarrow \mathrm{Lines}_{\mathbb C},
\]
with the orientation convention that a boundary partition vector
\(Z_{\partial,M}(A)\) lies in \(\mathcal L_M(A)\).  When fillings exist,
\(\mathcal L_M(A)\) may be constructed by the filling quotient described in
Section~\ref{sec:inflow-bulk-boundary-lines}; otherwise the functor
\(\mathcal L\) is taken as the boundary value of the invertible bulk.
\item \(Z_{\partial}\) is a relative \(D\)-dimensional QFT partition
assignment: for each closed \((M,A)\),
\[
  Z_{\partial,M}(A)\in \mathcal L_M(A),
\]
with covariance under background isomorphisms and with the usual gluing law
after pairing with compatible bulk states.
\end{enumerate}
A trivialization of the anomaly is a monoidal natural isomorphism
\(\mathcal L\simeq \mathbf 1\), possibly after tensoring with a local
counterterm invertible theory or with an inverse anomalous sector.  A
coordinate anomaly formula, such as a descent form or a finite cochain
cocycle, is a representative of this structure after choices of local frames
and local representatives have been made.
\end{definition}

This definition is the object controlled throughout the chapter.  The
connected Chern--Weil descent, the finite higher-form cochain models, the
Weyl-fermion index representative, and the noninvertible wall examples below
are different coordinate presentations of the same kind of background-line and
invertible-bulk structure.

\paragraph{Coordinate comparison.}
The anomaly line has three coordinate shadows used in different volumes.
In a finite-regulator chart such as the one used in
Chapter~\ref{ch:volume-ii-20}, a background transformation \(g:A\to A^g\)
is represented by the additive cochain
\begin{equation}
  C_\Lambda(g;A)
  =
  W_\Lambda[A^g]-W_\Lambda[A]
  \quad\in\mathbb R/\mathbb Z ,
  \qquad
  C_\Lambda(gh;A)
  =
  C_\Lambda(g;A)+C_\Lambda(h;A^g).
  \label{eq:inflow-finite-regulator-cochain-coordinate}
\end{equation}
After a local frame \(s_A\) for the anomaly line is chosen, the same
information is the multiplicative line cocycle
\begin{equation}
  \alpha_\Lambda(A,g)
  =
  \exp\!\bigl(2\pi\ii C_\Lambda(g;A)\bigr),
  \qquad
  \mathcal L(g)s_A=\alpha_\Lambda(A,g)s_{A^g}.
  \label{eq:inflow-cochain-to-line-cocycle-coordinate}
\end{equation}
Adding a local finite-regulator counterterm \(B_\Lambda(A)\) shifts the
effective-action representative by
\begin{equation}
  C_\Lambda(g;A)
  \longmapsto
  C_\Lambda(g;A)+B_\Lambda(A^g)-B_\Lambda(A).
  \label{eq:inflow-regulator-counterterm-coboundary-coordinate}
\end{equation}
Equivalently, moving that counterterm from the scalar coordinate into the line
frame uses the inverse frame change
\(s_A\mapsto\exp(-2\pi\ii B_\Lambda(A))s_A\), giving the same transformed
\(\alpha_\Lambda\).

The inflow bulk carries the inverse coordinate.  If a cylinder or mapping
cylinder realizes the transformation \(g\), then the bulk transition phase on
that bordism is
\begin{equation}
  \mathcal Z_{\rm bulk}([0,1]\times_g M)
  =
  \alpha_\Lambda(A,g)^{-1}
  =
  \exp\!\bigl(-2\pi\ii C_\Lambda(g;A)\bigr),
  \label{eq:inflow-bulk-inverse-line-coordinate}
\end{equation}
so that the boundary vector and the bulk state pair to a background-invariant
number.  In the free-fermion global-anomaly chapter the same line cocycle is
computed analytically by determinant-line transport,
\[
  \alpha_{\rm det}(A,g)
  =
  \operatorname{Hol}_g(\operatorname{Det}D)
  =
  \exp\!\left(-2\pi\ii\,\xi(B_{Y_g})\right),
\]
or by the corresponding Pfaffian sign in the real case.  Thus the finite
cochain, the inflow boundary line, and the determinant/Pfaffian holonomy are
not separate anomaly notions; they are regulator, invertible-bulk, and
global-analysis coordinates on the same obstruction to descent.  The analytic
theorems quoted in Chapter~\ref{ch:curved-eta-invariants-global-anomalies}
compute this coordinate for Dirac families, while the algebra above is the
QFT-facing descent and cancellation bookkeeping.

\section{Background Fields and Relative Partition Functions}

Let \(M\) be a closed \(D\)-manifold with the spacetime structure required by
the theory: orientation, spin structure, metric, conformal class, or another
declared structure.  Let
\[
  \mathcal B(M)
\]
be the background-field groupoid in the anomaly-inflow structure
\(\mathfrak A\) of Definition~\ref{def:anomaly-inflow-datum}.  An object \(A\in
\mathcal B(M)\) may contain ordinary gauge fields, higher-form gauge fields,
metric data, spin connections, and background defect data.  A morphism
\[
  u:A\longrightarrow A'
\]
is a background-field isomorphism, for example a gauge transformation,
higher-form gauge transformation, or diffeomorphism preserving the declared
spacetime structure.

For a nonanomalous theory the partition function is a function on isomorphism
classes:
\[
  Z_M(A')=Z_M(A)
  \quad\text{whenever }A\simeq A'.
\]
For an anomalous theory the correct object is a complex line over each
background.  Write
\[
  \mathcal L_M:\mathcal B(M)\longrightarrow
  \mathrm{Lines}_{\mathbb C}
\]
for a functor from the background groupoid to the groupoid of one-dimensional
complex vector spaces and nonzero linear maps.  The anomalous partition
function is a section
\[
  Z_M(A)\in \mathcal L_M(A).
\]
For a morphism \(u:A\to A'\), covariance means
\[
  Z_M(A')=\mathcal L_M(u)Z_M(A),
\]
not equality of complex numbers.  The functor \(\mathcal L_M\) is the anomaly
line.

Choosing a nonzero vector \(s_A\in\mathcal L_M(A)\) for every object \(A\)
trivializes the line as a collection of vector spaces, but generally not as a
functor.  In such a trivialization,
\[
  Z_M(A)=z_M(A)s_A,\qquad z_M(A)\in\mathbb C,
\]
and every morphism \(u:A\to A'\) acts by a nonzero phase or scalar
\[
  \mathcal L_M(u)s_A=\alpha_M(u;A)s_{A'}.
\]
Thus
\[
  z_M(A')=\alpha_M(u;A)z_M(A).
\]
The scalar \(\alpha_M\) is not itself the anomaly; it is the anomaly written
in a chosen local trivialization.

\begin{remark}[Functorial cocycle condition]
\label{rem:anomaly-line-functorial-cocycle}
Let \(u:A\to A'\) and \(v:A'\to A''\) be composable morphisms in
\(\mathcal B(M)\).  In any choice of nonzero local frame \(s_A\) for the
anomaly line,
\[
  \alpha_M(vu;A)
  =
  \alpha_M(v;A')\alpha_M(u;A).
\]
If the frame is changed by
\[
  s_A'=c(A)s_A,\qquad c(A)\in\mathbb C^\times,
\]
then
\[
  \alpha'_M(u;A)
  =
  c(A')^{-1}\alpha_M(u;A)c(A).
\]
Functoriality gives
\[
  \mathcal L_M(vu)
  =
  \mathcal L_M(v)\mathcal L_M(u).
\]
Applying both sides to \(s_A\) gives
\[
  \alpha_M(vu;A)s_{A''}
  =
  \alpha_M(u;A)\alpha_M(v;A')s_{A''},
\]
which is the stated cocycle identity.  If \(s'_A=c(A)s_A\), then
\[
  \mathcal L_M(u)s'_A
  =
  c(A)\alpha_M(u;A)s_{A'}
  =
  c(A)\alpha_M(u;A)c(A')^{-1}s'_{A'}.
\]
This is exactly the displayed transformation law.
\end{remark}

A local counterterm is a choice of local functional \(K_M(A)\) whose
exponential changes the frame:
\[
  s_A\longmapsto
  \exp(2\pi\ii K_M(A))s_A.
\]
It changes the representative cocycle by a coboundary.  Therefore an anomaly
class is not a particular variation formula; it is the isomorphism class of
the line functor together with its locality and gluing structure.

\section{Invertible Bulk Theory and Boundary Lines}
\label{sec:inflow-bulk-boundary-lines}

For the same structure \(\mathfrak A\), the category
\(\mathrm{Bord}_{D+1}^{\mathcal B}\) has closed \(D\)-manifolds with
background fields as objects and \((D+1)\)-dimensional bordisms with
compatible background fields as morphisms.  Its invertible bulk entry is a
symmetric monoidal functor
\[
  \mathcal Z_{\mathrm{bulk}}:
  \mathrm{Bord}_{D+1}^{\mathcal B}
  \longrightarrow
  \mathrm{Lines}_{\mathbb C}
\]
such that every closed \((D+1)\)-manifold \(X\) is assigned a nonzero complex
number and every \(D\)-manifold \(M\) is assigned a line.  If
\[
  X=X_1\cup_M X_2,
\]
then the value on \(X\) is obtained by pairing the line assigned to \(M\) from
\(X_1\) with the dual line assigned from \(X_2\).  This gluing axiom is the
mathematical content behind the word ``inflow''.

When the bulk is represented locally by a \((D+1)\)-cocycle
\[
  I_{D+1}(\widetilde A)\in C^{D+1}(X,\mathbb R/\mathbb Z)
\]
or by a differential-cocycle representative, its closed-manifold value is
\[
  \mathcal Z_{\mathrm{bulk}}(X,\widetilde A)
  =
  \exp\left(2\pi\ii\int_X I_{D+1}(\widetilde A)\right).
\]
The periods are integral in the normalization used, so changing a local
representative by a globally defined coboundary leaves the closed value
unchanged.

There is a concrete construction of the boundary anomaly line from the bulk.
Fix \(A\in\mathcal B(M)\).  A vector in the line \(\mathcal L_M(A)\) may be
represented by a pair
\[
  (X,\widetilde A),
  \qquad
  \partial X=M,\quad
  \widetilde A|_M=A,
\]
together with a complex coefficient.  Two representatives
\((X,\widetilde A,z)\) and \((X',\widetilde A',z')\) are identified if
\[
  z'
  =
  z\,
  \mathcal Z_{\mathrm{bulk}}
  \bigl((-X)\cup_M X',\,(-\widetilde A)\cup_A\widetilde A'\bigr).
\]
This quotient is one-dimensional whenever the background admits the relevant
extensions and the usual bordism-locality assumptions hold.  A relative
boundary theory assigns a vector in this line.  Multiplying the boundary
vector by the bulk number for a filling gives an ordinary complex number, but
only after the filling has been chosen.

\paragraph{Bulk gluing and the anomaly line.}
The construction above defines a line functor over the background groupoid of
\(M\).  If the bulk invertible theory is trivializable, the line functor is
trivializable; if a boundary theory has partition vectors in this line, then
the combined bulk-boundary partition function on a closed glued manifold is
an ordinary complex number invariant under background isomorphisms.
The mathematical content rests on the functorial gluing law of the bulk
invertible theory, the existence of the allowed fillings or the corresponding
colimit over fillings, and the lifting of background isomorphisms to the
filling data.  Under these inputs, transitivity of the equivalence relation
follows from the closed-manifold
gluing axiom:
\[
  \mathcal Z((-X)\cup_M X'')=
  \mathcal Z((-X)\cup_M X')\,
  \mathcal Z((-X')\cup_M X'')
\]
with the common boundary \(X'\) paired with its orientation reverse.  A
background isomorphism \(u:A\to A'\) maps a filling of \(A\) to the same
filling with boundary background identified by \(u\), hence gives a linear
map between the corresponding quotient lines.  Functoriality of these maps is
again the gluing axiom.  A trivialization of the bulk functor supplies
canonical vectors in all boundary lines, and therefore trivializes the line
functor.  In the absence of such a trivialization, the boundary partition
function is a vector in the anomaly line, and pairing with a compatible bulk
filling produces a number.  Gauge covariance of the number follows because
the bulk and boundary line transformations are inverse under this pairing.

Figure~\ref{fig:inflow-boundary-line-gluing} records the gluing statement for
anomaly inflow: a boundary partition vector pairs with a compatible bulk
filling to give a gauge-covariant number.

\begin{figure}[H]
\centering
\begin{tikzpicture}[scale=0.95, every node/.style={font=\small}]
  \fill[gray!10] (-3.1,-1.1) rectangle (-0.35,1.1);
  \draw[thick] (-0.35,-1.1) -- (-0.35,1.1);
  \node at (-1.75,0.15) {\((X,\widetilde A)\)};
  \node at (-0.15,1.32) {\(M,A\)};
  \draw[->,thick] (0.35,0) -- (1.25,0);
  \fill[gray!10] (1.9,-1.1) rectangle (4.65,1.1);
  \draw[thick] (1.9,-1.1) -- (1.9,1.1);
  \node at (3.25,0.15) {\((X',\widetilde A')\)};
  \node at (1.7,1.32) {\(-M,A\)};
  \node at (0.8,-0.55) {glue};
  \draw[rounded corners, thick] (-3.1,-1.1) -- (-0.35,-1.1) -- (1.9,-1.1)
    -- (4.65,-1.1) -- (4.65,1.1) -- (1.9,1.1) -- (-0.35,1.1) -- (-3.1,1.1)
    -- cycle;
  \node at (0.8,-1.45) {closed bulk phase relates boundary-line frames};
\end{tikzpicture}
\caption{A boundary anomaly line is obtained by comparing fillings.  Gluing
two fillings along their common boundary produces the closed invertible-bulk
phase that changes the frame.}
\label{fig:inflow-boundary-line-gluing}
\end{figure}

\section{Descent for Connected Symmetries}

Let \(A\) be a connection for a connected Lie symmetry and let
\[
  F=dA+A\wedge A
\]
be its curvature.  Let \(P_{D+2}(F)\) be a closed invariant polynomial of
degree \(D+2\), normalized so that its integral on every closed
\((D+2)\)-manifold with the corresponding bundle is integral.  Locally one
can choose a Chern--Simons form \(I_{D+1}^{(0)}(A)\) satisfying
\[
  dI_{D+1}^{(0)}(A)=P_{D+2}(F).
\]
For the common case \(D=2n\) and
\[
  P_{2n+2}(F)=\langle F^{n+1}\rangle ,
\]
where \(\langle\cdots\rangle\) is an invariant symmetric
\((n+1)\)-linear form on the Lie algebra, this local primitive is
constructed by a path of connections.  Put
\[
  A_t=tA,\qquad
  F_t=dA_t+A_t\wedge A_t=t\,dA+t^2A\wedge A .
\]
Define
\begin{equation}
  I_{2n+1}^{(0)}(A)
  =
  (n+1)\int_0^1dt\,\langle A\,F_t^n\rangle .
  \label{eq:universal-cs-transgression}
\end{equation}

\paragraph{Chern--Weil transgression.}
The form \eqref{eq:universal-cs-transgression} satisfies
\[
  d I_{2n+1}^{(0)}(A)=\langle F^{n+1}\rangle .
\]

Let \(D_t=d+[A_t,\cdot]\) be the covariant exterior derivative for \(A_t\).
Since \(\dot A_t=A\), the curvature variation is
\[
  \dot F_t=D_t\dot A_t=D_tA .
\]
Using invariance of \(\langle\cdots\rangle\) and the Bianchi identity
\(D_tF_t=0\),
\[
  \frac{d}{dt}\langle F_t^{n+1}\rangle
  =
  (n+1)\langle D_tA\,F_t^n\rangle
  =
  (n+1)d\langle A\,F_t^n\rangle .
\]
Integrating this identity from \(t=0\) to \(t=1\) gives
\[
  \langle F^{n+1}\rangle-\langle F_0^{n+1}\rangle
  =
  d\left((n+1)\int_0^1dt\,\langle A\,F_t^n\rangle\right).
\]
Because \(F_0=0\), the left side is \(\langle F^{n+1}\rangle\), proving the
claim.

In the Abelian case \(F=dA\) and \eqref{eq:universal-cs-transgression}
reduces to
\[
  I_{2n+1}^{(0)}(A)=A\wedge F^n .
\]
For a gauge variation \(\delta_\lambda A=d\lambda\),
\[
  \delta_\lambda I_{2n+1}^{(0)}
  =
  d\lambda\wedge F^n
  =
  d(\lambda F^n),
  \qquad dF=0,
\]
so \(I_{2n}^{(1)}(\lambda,A)=\lambda F^n\) is an explicit descent
representative in this normalization.
For an infinitesimal gauge parameter \(\alpha\),
\[
  \delta_\alpha A=D_A\alpha=d\alpha+[A,\alpha].
\]
Since \(P_{D+2}(F)\) is gauge invariant,
\[
  d\,\delta_\alpha I_{D+1}^{(0)}(A)=0.
\]
Locally, and globally after passing to the descent complex, this gives a
\(D\)-form \(I_D^{(1)}(\alpha,A)\) such that
\[
  \delta_\alpha I_{D+1}^{(0)}(A)=dI_D^{(1)}(\alpha,A).
\]
Therefore, for \(X^{D+1}\) with boundary \(M\),
\[
  \delta_\alpha
  2\pi\ii\int_X I_{D+1}^{(0)}
  =
  2\pi\ii\int_M I_D^{(1)}(\alpha,A).
\]
The boundary generating functional must transform by the negative of this
phase for the combined bulk-boundary object to be gauge invariant.

\begin{proposition}[Wess--Zumino consistency from descent]
\label{prop:wess-zumino-consistency-from-descent}
Let \(M\) be a closed \(D\)-manifold and let
\[
  \delta_\alpha W_M[A]
  =
  2\pi\ii\int_M I_D^{(1)}(\alpha,A)
\]
be obtained from a gauge-invariant polynomial by descent.  Then
\[
  \delta_\alpha\delta_\beta W_M
  -
  \delta_\beta\delta_\alpha W_M
  =
  \delta_{[\alpha,\beta]}W_M .
\]
\end{proposition}

\begin{proof}
Introduce a ghost \(c\) valued in the Lie algebra and the BRST differential
\(\mathsf s A=D_Ac\), \(\mathsf sc=-\frac12[c,c]\).  Gauge invariance of
\(P(F)\) and the Chern--Weil transgression give a local descent tower
\[
  \mathsf s I_{D+1}^{(0)}+dI_D^{(1)}=0,
  \qquad
  \mathsf s I_D^{(1)}+dI_{D-1}^{(2)}=0,
\]
with the sign convention chosen so that the component of the first equation
linear in a ghost value \(\alpha\) is
\[
  \delta_\alpha I_{D+1}^{(0)}(A)=dI_D^{(1)}(\alpha,A).
\]
We now extract the ghost-number-two coefficient explicitly.  Introduce two
odd scalar tags \(\varepsilon,\eta\), put
\[
  c=\varepsilon\alpha+\eta\beta,
  \qquad
  \varepsilon\eta=-\eta\varepsilon,\qquad
  \varepsilon^2=\eta^2=0.
\]
The BRST rule gives
\[
  [c,c]=2\varepsilon\eta[\alpha,\beta],
  \qquad
  \mathsf sc=-\varepsilon\eta[\alpha,\beta],
\]
while
\[
  \mathsf sA
  =
  \varepsilon D_A\alpha+\eta D_A\beta .
\]
Since \(I_D^{(1)}(c,A)\) is linear in \(c\), the coefficient of
\(\varepsilon\eta\) in \(\mathsf s I_D^{(1)}(c,A)\) is
\begin{equation}
  \delta_\alpha I_D^{(1)}(\beta,A)
  -
  \delta_\beta I_D^{(1)}(\alpha,A)
  -
  I_D^{(1)}([\alpha,\beta],A).
  \label{eq:wz-ce-cochain-from-ghost-two}
\end{equation}
The first two terms come from the variation of \(A\); the relative minus sign
is the Koszul sign obtained when the odd tag of the varied parameter crosses
the odd BRST differential.  The last term comes from
\(\mathsf sc=-\varepsilon\eta[\alpha,\beta]\).  Thus the second descent
equation gives
\begin{equation}
  \delta_\alpha I_D^{(1)}(\beta,A)
  -
  \delta_\beta I_D^{(1)}(\alpha,A)
  -
  I_D^{(1)}([\alpha,\beta],A)
  =
  -d\,I_{D-1}^{(2)}(\alpha,\beta,A),
  \label{eq:wz-ce-coboundary-is-total-derivative}
\end{equation}
where \(I_{D-1}^{(2)}(\alpha,\beta,A)\) denotes the
\(\varepsilon\eta\)-coefficient of \(I_{D-1}^{(2)}(c,A)\).

Integrating \eqref{eq:wz-ce-coboundary-is-total-derivative} over closed \(M\)
annihilates the right side.  Multiplication by \(2\pi\ii\) gives
\[
  \delta_\alpha\delta_\beta W_M
  -
  \delta_\beta\delta_\alpha W_M
  -
  \delta_{[\alpha,\beta]}W_M
  =
  0.
\]
This is the Chevalley--Eilenberg cocycle condition for the local anomaly
one-cochain
\[
  \mathcal A_M(\alpha;A)
  =
  2\pi\ii\int_M I_D^{(1)}(\alpha,A).
\]
A change of descent representative by a local counterterm adds a
Chevalley--Eilenberg coboundary, so the identity survives the representative
change.  If \(M\) has a boundary, the right side of
\eqref{eq:wz-ce-coboundary-is-total-derivative} is the boundary descent term
that must be included in the corresponding relative anomaly-line structure.
\end{proof}

\begin{remark}[Cocycle ambiguity and local counterterms]
\label{rem:inflow-cocycle-counterterm}
Changing \(I_{D+1}^{(0)}\) by
\[
  I_{D+1}^{(0)}\longmapsto I_{D+1}^{(0)}+dK_D(A)
\]
changes the boundary anomaly by the variation of the local
\(D\)-dimensional functional
\[
  2\pi\ii\int_M K_D(A).
\]
Thus the anomaly invariant is the cohomology class of the bulk invertible
theory, while a representative anomaly current depends on a boundary
counterterm coordinate.

On a manifold with boundary,
\[
  \int_X dK_D=\int_M K_D .
\]
Under a background gauge transformation, the additional boundary variation is
exactly the variation of this \(D\)-dimensional local functional.  Hence two
bulk cocycles differing by a coboundary define anomaly representatives that
are related by a local counterterm.  The same statement in the anomaly-line
language is the frame change in
Remark~\ref{rem:anomaly-line-functorial-cocycle}.
\end{remark}

\section{Finite Higher-Form Inflow}

The connected descent formalism is not the whole theory of anomalies.  Finite
symmetries and higher-form symmetries are naturally described by cochains.
Let \(A\) be a finite abelian group and let a \(p\)-form background on a
\((D+1)\)-manifold \(X\) be
\[
  B\in Z^{p+1}(X,A),
  \qquad
  B\sim B+\delta\Lambda .
\]
An invertible finite response is represented by a cocycle
\[
  \omega(B)\in Z^{D+1}(X,\mathbb R/\mathbb Z)
\]
natural under subdivision and pullback.  Its value is
\[
  \exp\left(2\pi\ii\int_X\omega(B)\right).
\]
If \(X\) is closed and \(\omega\) is a cocycle on the classifying space
\(B^{p+1}A\), gauge invariance follows by the same evaluation-on-cycles
argument as Remark~\ref{rem:discrete-theta-coboundary-invariance}.  If
\(X\) has boundary, gauge variation leaves a boundary cochain.  The boundary
QFT is then a section of the corresponding finite anomaly line.

For one-form symmetries in four dimensions the bulk is five-dimensional.  A
particularly useful local model is the finite Heisenberg response with
\[
  b,c\in C^2(X,\mathbb Z_N)
\]
and
\[
  S_H[X;b,c]
  =
  \exp\left(
    \frac{2\pi\ii}{N}\int_X b\smile\delta c
  \right).
\]
Here \(b\) and \(c\) are the two finite two-form fields paired by the
\(\mathbb Z_N\) Pontryagin duality.  On a closed \(X\) the action is invariant
under
\[
  b\mapsto b+\delta\lambda,\qquad c\mapsto c+\delta\kappa,
  \qquad
  \lambda,\kappa\in C^1(X,\mathbb Z_N).
\]
On a manifold with boundary \(M\), the first transformation changes the
exponent by
\[
  \frac{2\pi\ii}{N}\int_M \lambda\smile\delta c.
\]
This is an anomaly for a boundary condition that keeps \(c|_M\) as a
background.  Different boundary conditions correspond to different choices of
which finite field is fixed, which is summed, and which boundary counterterm
has been attached.  This is the cochain model behind one-form gauging,
duality interfaces, and the symmetry-TQFT discussion in
Chapter~\ref{ch:global-higher-group-symmetry-tft}.

\begin{remark}[Finite bulk variation]
\label{rem:finite-higher-form-bulk-variation}
For the five-dimensional action
\[
  \frac{2\pi\ii}{N}\int_X b\smile\delta c,
\]
the variation under \(b\mapsto b+\delta\lambda\) is trivial on closed \(X\)
and equals
\[
  \frac{2\pi\ii}{N}\int_{\partial X}\lambda\smile\delta c
\]
when \(X\) has boundary.  The variation under
\(c\mapsto c+\delta\kappa\) is zero.

Since \(\delta^2=0\), changing \(c\) by \(\delta\kappa\) does not change
\(\delta c\).  Under \(b\mapsto b+\delta\lambda\), the exponent changes by
\[
  \frac{2\pi\ii}{N}\int_X \delta\lambda\smile\delta c.
\]
The cup-product differential satisfies
\[
  \delta(\lambda\smile\delta c)
  =
  \delta\lambda\smile\delta c
  -\lambda\smile\delta^2c
  =
  \delta\lambda\smile\delta c,
\]
because \(\deg\lambda=1\).  Evaluation of a coboundary on a closed
fundamental cycle is zero.  On a manifold with boundary, the cochain Stokes
formula gives the displayed boundary term.
\end{remark}

\section{Weyl-Fermion Inflow}
\label{sec:four-dimensional-weyl-five-dimensional-inflow}

The local descent formalism of Chapter~\ref{ch:volume-ii-20} becomes an
invertible five-dimensional theory after the anomaly polynomial is placed in
its index normalization.  Let \(X\) be an oriented five-manifold with boundary
\[
  M=\partial X.
\]
Let \(A_X\) be a background connection on \(X\), and let \(R_X\) be the
curvature of a Riemannian connection on \(TX\).  For one left-handed
four-dimensional Weyl fermion in a representation \(R\), the
index-normalized six-form is
\begin{equation}
  \mathcal I_6(R)
  =
  \left[
    \widehat A(TX)\operatorname{ch}_{R}(F_X)
  \right]_6 ,
  \label{eq:volix-index-normalized-six-form}
\end{equation}
with the notation of
Section~\ref{sec:index-normalized-anomaly-polynomials}.  Choose locally a
five-form \(\mathcal I_5^{(0)}\) satisfying
\[
  d\mathcal I_5^{(0)}=\mathcal I_6,
  \qquad
  \delta_\lambda\mathcal I_5^{(0)}
  =
  d\mathcal I_4^{(1)}(\lambda).
\]
The five-dimensional response
\begin{equation}
  W_X^{\rm inv}[A_X,g_X]
  =
  -2\pi\ii\int_X \mathcal I_5^{(0)}
  \label{eq:five-dimensional-weyl-inflow-action}
\end{equation}
has boundary variation
\[
  \delta_\lambda W_X^{\rm inv}
  =
  -2\pi\ii\int_M\mathcal I_4^{(1)}(\lambda).
\]
It cancels a boundary left-handed determinant with
\[
  \delta_\lambda W_M
  =
  +2\pi\ii\int_M\mathcal I_4^{(1)}(\lambda).
\]
The sign in \eqref{eq:five-dimensional-weyl-inflow-action} is a
boundary-condition choice: reversing the orientation of \(X\), or using a
right-handed boundary fermion, reverses the phase.

\paragraph{Coefficient of abelian inflow.}
Let the boundary field be one left-handed Weyl fermion of integer \(U(1)\)
charge \(q\).  Let \(F=dA\), and normalize the first Pontryagin form by
\[
  p_1(TX)
  =
  -\frac1{8\pi^2}
  \operatorname{tr}_{TX}(\mathsf R^2).
\]
Then a local representative of the inflow action is
\begin{equation}
  W_X^{\rm inv}
  =
  -2\pi\ii
  \int_X
  \left[
    \frac{q^3}{6(2\pi)^3} A F^2
    -
    \frac{q}{24(2\pi)} A\,p_1(TX)
  \right],
  \label{eq:abelian-weyl-inflow-action}
\end{equation}
and under \(A\mapsto A+d\lambda\),
\begin{equation}
  \delta_\lambda W_X^{\rm inv}
  =
  -2\pi\ii
  \int_M
  \lambda
  \left[
    \frac{q^3}{6(2\pi)^3}F^2
    -
    \frac{q}{24(2\pi)}p_1(TM)
  \right].
  \label{eq:abelian-weyl-inflow-variation}
\end{equation}

For \(U(1)\), the index-normalized polynomial is
\[
  \mathcal I_6(q)
  =
  \frac{q^3}{6(2\pi)^3}F^3
  -
  \frac{q}{24(2\pi)}p_1(TX)F .
\]
Because \(dF=0\) and \(dp_1=0\) for the Chern--Weil representative, a
primitive is the bracketed five-form in
\eqref{eq:abelian-weyl-inflow-action}.  Varying \(A\) gives
\[
  \delta_\lambda
  \left[
    \frac{q^3}{6(2\pi)^3}AF^2
    -
    \frac{q}{24(2\pi)}A\,p_1
  \right]
  =
  d\lambda
  \left[
    \frac{q^3}{6(2\pi)^3}F^2
    -
    \frac{q}{24(2\pi)}p_1
  \right].
\]
The expression in square brackets is closed, hence this variation is the
exterior derivative of
\[
  \lambda
  \left[
    \frac{q^3}{6(2\pi)^3}F^2
    -
    \frac{q}{24(2\pi)}p_1
  \right].
\]
Stokes' theorem gives \eqref{eq:abelian-weyl-inflow-variation} with the
boundary orientation \(M=\partial X\).

For several left-handed Weyl fermions, the local \(U(1)^3\) and mixed
gravitational-\(U(1)\) anomaly coefficients are
\[
  \sum_i q_i^3,\qquad \sum_i q_i .
\]
If the \(U(1)\) symmetry is dynamical, the anomaly line for the dynamical
gauge group must be canonically trivial.  In the absence of a four-dimensional
local counterterm that shifts these representatives, the local conditions are
\[
  \sum_i q_i^3=0,\qquad \sum_i q_i=0,
\]
together with the mixed nonabelian-\(U(1)\) conditions when other gauge
factors are present.  If the \(U(1)\) symmetry is a background global
symmetry, the same polynomial is retained as the response that must be
matched by any infrared description.  Local cancellation does not by itself
settle possible global anomalies; those are statements about the full
invertible theory on the relevant bordism category.

\section{From Inflow to Noninvertible Chiral Defects}
\label{sec:inflow-to-noninvertible-chiral-defects}

Inflow also defines certain noninvertible codimension-one operators.  Start
with a transformation \(g\) of a \(D\)-dimensional QFT.  Suppose that, in the
presence of background fields \(A\), implementing \(g\) across a hypersurface
\[
  \Sigma\subset M
\]
leaves an anomaly line described by an invertible \((D+1)\)-dimensional
response \(\mathcal Z_{\rm anom}(g;A)\).  The bare wall for \(g\) is then not
an ordinary topological symmetry wall.  It is a relative theory for the
restriction of the anomaly to a small tubular neighborhood of \(\Sigma\).

Let \(\mathcal D_g\) be the transformation wall.  A wall theory
\(\mathcal T_\Sigma\) produces a genuine defect precisely when there is a
trivialization
\[
  \mathcal L_{\mathcal D_g}\otimes
  \mathcal L_{\mathcal T_\Sigma}
  \simeq
  \mathbf 1
\]
of anomaly lines over the wall-background groupoid.  The honest defect is
\[
  D_g=\mathcal D_g\otimes\mathcal T_\Sigma .
\]
If \(\mathcal T_\Sigma\) is obtained by gauging a finite higher-form symmetry
on the wall, fusion of \(D_g\) with its orientation reverse contains a sum
over the gauged sectors.  That fusion is not a single copy of the identity
defect, and \(D_g\) is noninvertible.

\begin{remark}[Wall trivialization criterion]
\label{rem:wall-trivialization-criterion}
Assume the transformation wall \(\mathcal D_g\) and the auxiliary wall theory
\(\mathcal T_\Sigma\) are topological away from their anomaly lines.  The
tensor product \(D_g=\mathcal D_g\otimes\mathcal T_\Sigma\) is a genuine
topological defect exactly when the tensor product of their anomaly-line
functors is trivializable.  If the trivializing wall theory contains a
nontrivial finite gauge sum, then \(D_g\otimes D_g^\vee\) contains the
corresponding condensation projector and need not be the identity defect.

Topological invariance of a defect means that its correlation functional is a
number invariant under isotopy, with no uncancelled dependence on a choice of
background-field frame.  The product wall has anomaly line
\(\mathcal L_{\mathcal D_g}\otimes\mathcal L_{\mathcal T_\Sigma}\).  A
trivialization of this line converts its partition vector into an ordinary
number and removes the frame dependence, so the wall is genuine.  Conversely,
if the product line is nontrivial, changing the background by a morphism in
the wall groupoid changes the wall vector by a nontrivial line map; no
ordinary topological defect has this transformation law.  When
\(\mathcal T_\Sigma\) is a finite gauge theory, fusing with the orientation
reverse performs the finite gauge sum on the wall twice with the dual
orientation.  The result is the projector onto invariant wall sectors, not
necessarily the invisible wall.  This is the categorical source of
noninvertibility.
\end{remark}

\begin{example}[Axial defects in compact QED]
\label{ex:qed-noninvertible-defect-from-inflow}
Consider compact four-dimensional QED with a unit-charge massless Dirac
fermion and with no dynamical monopoles.  The axial rotation by angle
\(\alpha\) changes the effective action by the abelian anomaly
\[
  \delta_\alpha W
  =
  \ii\alpha\,\frac1{4\pi^2}
  \int_M F\wedge F
\]
in the convention of \eqref{eq:abelian-axial-anomaly}.  For
\(\alpha=2\pi/N\), the phase is sensitive to the magnetic one-form background
when the theory is placed in sectors with fractional \(F/2\pi\) induced by
that background.  The codimension-one axial wall therefore carries a mixed
anomaly with the magnetic one-form symmetry.  Attaching a three-dimensional
finite gauge theory whose Pontryagin-square variation cancels this mixed
phase gives the defects \(D_{N,p}\) constructed in
Chapter~\ref{ch:global-higher-group-symmetry-tft}.  The role of the
three-dimensional theory is fixed by the inflow equation above: it is the
boundary trivialization of the anomaly line of the axial wall, not an
additional optional decoration.
\end{example}

\section{One-Form Symmetry Example}

Let a four-dimensional theory have a finite abelian one-form symmetry \(A\).
A background is a two-cocycle
\[
  B\in Z^2(M,A).
\]
An anomaly involving this symmetry is represented by a five-dimensional
cohomology class
\[
  \omega\in H^5(B^2A,U(1))
\]
or, with spacetime structure included, by the corresponding generalized
cohomology or bordism invariant.  For \(A=\mathbb Z_N\), expressions built
from cup products and the Pontryagin square give concrete representatives
after the manifold orientation and spin structure have been specified.  The
boundary theory then has line operators whose fusion and braiding with
surface defects are acted on by the anomaly line; changing a
four-dimensional discrete theta term shifts the representative but not the
inflow class unless the bulk cohomology class changes.

The finite one-form anomaly is detected operationally as follows.  Couple the
theory to \(B\).  If a one-form gauge transformation
\[
  B\mapsto B+\delta\Lambda
\]
changes the generating functional by
\[
  Z_M[B+\delta\Lambda]
  =
  \exp\left(2\pi\ii\int_M \eta(B,\Lambda)\right)Z_M[B],
\]
then \(\eta\) is a representative of the anomaly-line functor.  If there is a
local four-dimensional cochain \(K(B)\) such that
\[
  \eta(B,\Lambda)=K(B+\delta\Lambda)-K(B)
  \quad\text{mod }\mathbb Z,
\]
the anomaly is removed by the counterterm
\(\exp(-2\pi\ii\int_MK(B))\).  If no such local cochain exists in the chosen
background-field theory, gauging the one-form symmetry is obstructed unless
another sector supplies the inverse anomaly.

\section{Gauging Criterion}

To gauge a global symmetry, one must sum or integrate over its backgrounds.
For a finite symmetry this means a groupoid cardinality weighted sum; for a
continuous symmetry it means a regularized path integral over connections and
gauge transformations.  In either case the integrand must be an ordinary
number on the background groupoid.  If the partition function is a section of
a nontrivial anomaly line, there is no canonical summand to add.

In the invertible-language formulation the criterion is
\[
  \mathcal Z_{\mathrm{bulk}}\otimes
  \mathcal Z_{\mathrm{counter}}
  \simeq 1
\]
as an invertible theory over the background-field bordism category.  A local
counterterm supplies such a trivialization exactly when the anomaly class is
a coboundary in the chosen background-field cohomology theory.  A second QFT
sector can also supply the inverse line; this is anomaly cancellation by
tensor product.

\begin{remark}[Gauging requires a trivialized anomaly line]
\label{rem:gauging-requires-trivialized-anomaly-line}
Let a finite global symmetry have background groupoid \(\mathcal B(M)\).
The gauged partition function
\[
  Z_{\mathrm{gauged}}(M)
  =
  \sum_{[A]\in\pi_0\mathcal B(M)}
  \frac{1}{|\operatorname{Aut}(A)|}\,Z_M(A)
\]
is canonically defined only after the anomaly line \(\mathcal L_M\) has been
trivialized compatibly with morphisms in \(\mathcal B(M)\).  If two
trivializations differ by a local counterterm, the two gauged theories differ
by that counterterm before summation.

The summation formula adds complex numbers.  If \(Z_M(A)\) lies in
\(\mathcal L_M(A)\), then terms belonging to different isomorphism classes
live in unrelated one-dimensional vector spaces.  A trivialization identifies
all these vectors with complex numbers.  Compatibility with morphisms is
required because the factor \(1/|\operatorname{Aut}(A)|\) divides by
automorphisms; if an automorphism acts nontrivially on the line, the would-be
summand is frame dependent.  A second trivialization differs from the first
by a nowhere-zero function on the background groupoid.  If this function is
the exponential of a local cochain, it is exactly a local counterterm, and it
weights each background sector before the groupoid sum is performed.
\end{remark}

\begin{openproblem}[Operator-level inflow]
\label{op:operator-level-inflow}
Develop a reconstruction theorem that starts from an anomalous local QFT,
constructs its anomaly line over background fields, and derives the action of
the inflow theory on extended operators, defect junctions, and Hilbert spaces
on manifolds with boundary.
\end{openproblem}
