\chapter{Higher-Group Symmetry And Symmetry TQFT}
\label{ch:global-higher-group-symmetry-tft}

\ConventionsMixed

Higher-form symmetries and ordinary symmetries may combine into a single
background-field structure.  Its gauge transformations and anomalies may be
linked by Postnikov classes, so the combined object carries more structure
than a direct product of groups.  This chapter defines higher-group
backgrounds and then explains the symmetry-TQFT organization of boundary
conditions and gauging operations.

\begin{hypothesis}[Higher-group symmetry presentation]
\label{hyp:higher-group-symmetry-tft-datum}
A \(D\)-dimensional QFT with finite two-group symmetry and symmetry-TQFT
presentation is specified, for the purposes of this chapter, by the tuple
\[
  \mathfrak S_{\mathcal Q,\mathbb G}
  =
  \bigl(
    \mathcal Q,\mathbb G,\mathcal B_{\mathbb G},
    \mathcal L_{\rm anom},
    \mathfrak T_{\rm sym},\mathfrak B_{\mathcal Q},
    \mathcal D_{\rm top},\tau
  \bigr).
\]
The entries are as follows.
\begin{enumerate}[label=\textnormal{(\roman*)}]
\item \(\mathcal Q\) is a \(D\)-dimensional local QFT in the sense used in the
  earlier foundational volumes.
\item \(\mathbb G=(G_0,A,\rho,\beta)\) is a finite two-group: \(G_0\) is a
  finite zero-form symmetry group, \(A\) is a finite abelian one-form symmetry
  group, \(\rho:G_0\to\operatorname{Aut}(A)\) is the action on one-form
  charges, and \(\beta\in H^3(BG_0,A_\rho)\) is the Postnikov class.
\item \(\mathcal B_{\mathbb G}(M)\) is the finite higher groupoid of
  backgrounds on a collared \(D\)-manifold \(M\).  In a triangulated chart its
  objects are pairs \((g,B)\) satisfying the twisted cocycle equation
  \(\delta_g B=\beta(g)\), and its higher morphisms are the zero-form,
  one-form, and gauge-for-gauge transformations preserving this equation.
\item \(\mathcal L_{\rm anom}\) is the anomaly-line functor over
  \(\mathcal B_{\mathbb G}\), equivalently the boundary value of an invertible
  \((D+1)\)-dimensional theory for the same background fields.
\item \(\mathfrak T_{\rm sym}\) is the \((D+1)\)-dimensional symmetry TQFT
  whose bulk fields are the \(\mathbb G\)-backgrounds and whose boundary
  conditions include background-fixing and summing/gauging boundaries.
\item \(\mathfrak B_{\mathcal Q}\) is the boundary condition of
  \(\mathfrak T_{\rm sym}\) representing the QFT \(\mathcal Q\).  Coupling
  \(\mathcal Q\) to a background means placing \(\mathfrak B_{\mathcal Q}\)
  opposite a background boundary of \(\mathfrak T_{\rm sym}\).
\item \(\mathcal D_{\rm top}\) is the corresponding topological defect system
  in \(\mathcal Q\): symmetry surfaces, walls, junction lines, endpoint
  operators, and their isotopy and Pachner-move identifications.
\item \(\tau\) is any chosen anomaly trivialization or discrete counterterm
  needed to turn a proposed gauging or condensation operation into an ordinary
  complex-valued path integral.
\end{enumerate}
All finite cochain calculations below are consequences of such a presentation
after a triangulated model for \(\mathcal B_{\mathbb G}\) has been chosen.  The
existence of \(\mathfrak T_{\rm sym}\), \(\mathfrak B_{\mathcal Q}\), and
\(\mathcal D_{\rm top}\) for an interacting continuum theory is a separate QFT
construction problem; see
Open Problem~\ref{op:higher-group-examples}.
\end{hypothesis}

\section{Two-Group Backgrounds}

Let \(\mathfrak S_{\mathcal Q,\mathbb G}\) be a presentation as in
Hypothesis~\ref{hyp:higher-group-symmetry-tft-datum}.  The two-group entry
\(\mathbb G=(G_0,A,\rho,\beta)\) consists of an ordinary finite internal
symmetry \(G_0\), an abelian one-form symmetry \(A\), an action of \(G_0\) on
\(A\), and a Postnikov class
\[
  \beta\in H^3(BG_0,A_\rho).
\]
On a triangulated spacetime \(M\), a background consists of a \(G_0\)-valued
one-cocycle \(g\) and an \(A\)-valued two-cochain \(B\) obeying
\[
  \delta_g B=\beta(g).
\]
Here \(\delta_g\) is the coboundary twisted by the \(G_0\)-action on \(A\).
Thus \(B\) is not closed when the ordinary background \(g\) has nontrivial
Postnikov obstruction.

\section{Gauge Transformations}

A \(G_0\) gauge transformation changes \(g\) by a coboundary.  The two-form
background \(B\) must transform at the same time so that
\(\delta_gB=\beta(g)\) remains valid.  A one-form gauge transformation with
\(A\)-valued one-cochain \(\lambda\) acts by
\[
  B\mapsto B+\delta_g\lambda .
\]
If \(\beta\ne0\), the ordinary and one-form transformations do not commute as
independent operations on the background fields.  The physical statement is
that charged line operators and local operators transform in a single
higher-group representation problem.

\section{Anomaly Functionals}

An anomaly of a two-group symmetry in \(D\) dimensions is an invertible
\((D+1)\)-dimensional action for \((g,B)\).  A basic term is
\[
  S_{\mathrm{anom}}[g,B]
  =
  2\pi i\int_X
  \langle B\smile \gamma(g)\rangle ,
\]
where \(\gamma(g)\) is a \(G_0\)-dependent cochain with values in the Pontryagin
dual of \(A\).  Gauge invariance on closed \(X\) imposes a cocycle condition
using \(\delta_gB=\beta(g)\).  If \(X\) has boundary, the boundary QFT
partition function is a section of the anomaly line over background fields.

\section{Symmetry TQFT}

For the presentation in Hypothesis~\ref{hyp:higher-group-symmetry-tft-datum}, the
object \(\mathfrak T_{\rm sym}\) packages background fields, gauging, and
duality defects in a \((D+1)\)-dimensional topological theory.  The physical
QFT \(\mathcal Q\) is represented by the boundary condition
\[
  \mathfrak B_{\mathcal Q}
\]
for the symmetry TQFT.  Coupling to a background corresponds to ending the
topological theory on a background boundary.  Gauging corresponds to replacing
that boundary condition by a summing boundary condition.

For an ordinary finite group \(G\), the symmetry TQFT is finite \(G\)-gauge
theory in one higher dimension.  For higher groups, the bulk field space is
the groupoid of pairs \((g,B)\) satisfying the Postnikov constraint.  The
fusion of topological defects in the QFT is inherited from the fusion of
topological boundary-changing operators in the bulk.

\section{\texorpdfstring{\(\mathbb Z_N\)}{ZN} One-Form Symmetry TQFT}
\label{sec:zn-one-form-symtft}

We now write the finite cochain model that will be used as the local model for
one-form gauging and condensation defects.  Let \(X\) be an oriented
triangulated five-manifold.  Let
\[
  b,c\in C^2(X,\mathbb Z_N)
\]
be two \(\mathbb Z_N\)-valued two-cochains.  Their gauge transformations are
\[
  b\mapsto b+\delta\lambda,\qquad
  c\mapsto c+\delta\kappa,
  \qquad
  \lambda,\kappa\in C^1(X,\mathbb Z_N).
\]
The finite Heisenberg, or \(BF\), symmetry TQFT has exponentiated action
\begin{equation}
  Z_H(X;b,c)
  =
  \exp\left(\frac{2\pi i}{N}\int_X b\smile\delta c\right).
  \label{eq:zn-one-form-symtft-action}
\end{equation}
The integral means evaluation of the resulting \(\mathbb Z_N\)-valued
five-cochain on the fundamental class and then embedding
\(\mathbb Z_N\hookrightarrow \mathbb R/\mathbb Z\).  Equivalently one chooses
integer lifts and keeps the result modulo \(N\); changing the lift changes the
exponent by an integral multiple of \(2\pi i\).

\begin{remark}[Closed-manifold gauge invariance]
\label{rem:zn-one-form-symtft-gauge-invariance}
If \(X\) is closed, \eqref{eq:zn-one-form-symtft-action} is invariant under
the gauge transformations above.

The transformation \(c\mapsto c+\delta\kappa\) does not change \(\delta c\),
because \(\delta^2=0\).  Under \(b\mapsto b+\delta\lambda\), the exponent
changes by
\[
  \frac{2\pi i}{N}\int_X \delta\lambda\smile\delta c .
\]
Since \(\deg\lambda=1\),
\[
  \delta(\lambda\smile\delta c)
  =
  \delta\lambda\smile\delta c
  -\lambda\smile\delta^2c
  =
  \delta\lambda\smile\delta c .
\]
Thus the change is the integral of a coboundary.  Evaluation of a coboundary on
the fundamental cycle of a closed triangulated manifold is zero modulo \(N\).
The exponent is therefore unchanged.
\end{remark}

If \(X\) has boundary \(M=\partial X\), the last computation leaves the
boundary term
\[
  \frac{2\pi i}{N}\int_M \lambda\smile\delta c .
\]
A boundary condition must therefore specify which combination of \(b\) and
\(c\) is fixed as a background and which is summed as a dynamical higher gauge
field.  A four-dimensional QFT \(\mathcal Q\) with a \(\mathbb Z_N\) one-form
symmetry is a boundary condition for this five-dimensional theory after the
choice of background boundary has been made.  Gauging the one-form symmetry is
the operation of replacing the boundary condition that fixes \(b\) by the
boundary condition that sums over \(b\) and fixes the dual background \(c\).

\begin{example}[Four-dimensional \(\mathbb Z_N\) line data]
\label{ex:zn-four-dimensional-line-data}
In a four-dimensional theory whose electric and magnetic one-form backgrounds
are both finite \(\mathbb Z_N\) backgrounds, a line charge may be labelled
abstractly by
\[
  (e,m)\in \mathbb Z_N\oplus\mathbb Z_N .
\]
The pairing controlling mutual locality is the finite Dirac pairing
\[
  \langle(e,m),(e',m')\rangle
  =
  em'-me'
  \quad\in\mathbb Z_N .
\]
A set of simultaneously genuine line operators is a subgroup on which this
pairing vanishes.  The five-dimensional action
\eqref{eq:zn-one-form-symtft-action} packages the two one-form backgrounds and
their finite symplectic pairing.  The wall that exchanges \(b\) and \(c\) acts
on line charges by
\[
  (e,m)\mapsto (m,-e),
\]
the finite analogue of the \(S\)-wall discussed in
Chapter~\ref{ch:global-duality-defects-gauging-orbifold-data}.  This statement
is about the line-charge lattice and its mutual-locality pairing; it is not an
assertion that arbitrary closed loops in four Euclidean dimensions have a
topological pairwise braiding invariant.
\end{example}

\section{Condensation And Gauging}

Let \(\mathcal C\) be the category or higher category of topological defects
describing a finite symmetry.  Gauging a sub-symmetry is condensation of a
commutative algebra object
\[
  A_{\mathrm{cond}}\in\mathcal C .
\]
The gauged theory has boundary condition obtained by allowing endpoints of
the condensed defects.  If the algebra object is obstructed by an anomaly,
the condensation exists only after coupling to an appropriate bulk or after
choosing a trivialization of the anomaly.

\begin{definition}[Triangulated higher-gauging defect]
\label{def:triangulated-higher-gauging-defect}
Let \(\mathcal Q\) be a four-dimensional QFT with an anomaly-free finite
\(\mathbb Z_N\) one-form symmetry.  Assume that for every closed oriented
surface \(\Sigma\) there are topological symmetry surfaces
\[
  U_n(\Sigma),\qquad n\in\mathbb Z_N,
\]
with junctions implementing \(U_nU_m\simeq U_{n+m}\), and assume these
surface networks are invariant under triangulation moves.  For a triangulated
oriented three-manifold \(Y\) supporting a codimension-one defect, define the
groupoid-normalized condensation insertion
\begin{equation}
  \mathfrak C(Y)
  =
  \frac{|C^0(Y,\mathbb Z_N)|}{|C^1(Y,\mathbb Z_N)|}
  \sum_{b\in Z^2(Y,\mathbb Z_N)}
  \mathsf U_b(Y).
  \label{eq:triangulated-one-form-condensation-defect}
\end{equation}
Here \(b\) is a closed two-cochain on \(Y\).  The symbol
\(\mathsf U_b(Y)\) denotes the topological wall network obtained from the
Poincare-dual one-cycle of \(b\): locally one thickens each dual edge in a
small normal collar of \(Y\) and inserts the corresponding one-form symmetry
surface segment.  The condition \(\delta b=0\) is the conservation law at the
vertices of the dual one-cycle, equivalently the absence of an uncancelled
endpoint for the wall network.  The factor
\(|C^0|/|C^1|\) is the finite groupoid volume of the two-form gauge
parameter \(C^1(Y,\mathbb Z_N)\) together with its degree-zero
gauge-for-gauge redundancy.  Omitting the \(|C^0|\) factor gives a
cellularly dependent representative; changing to a reduced convention divides
by the constant factor \(|H^0(Y,\mathbb Z_N)|\) on each connected component
and changes the defect by an invertible normalization convention rather than
by a dynamical QFT operation.
\end{definition}

\paragraph{Finite two-form gauge groupoid behind the normalization.}
The coefficient in
\eqref{eq:triangulated-one-form-condensation-defect} is the homotopy
cardinality of the finite two-form gauge groupoid, written before quotienting
the field variables.  Let
\(\mathcal Z_N^2(Y)\) be the finite \(2\)-groupoid whose objects are closed
two-cochains
\[
  b\in Z^2(Y,\mathbb Z_N),
\]
whose \(1\)-morphisms are one-cochains
\(\lambda\in C^1(Y,\mathbb Z_N)\) acting by
\[
  b\longmapsto b+\delta\lambda,
\]
and whose \(2\)-morphisms are zero-cochains
\(\xi\in C^0(Y,\mathbb Z_N)\) acting on gauge parameters by
\[
  \lambda\longmapsto \lambda+\delta\xi .
\]
For a closed connected component this is the finite model of a
\(\mathbb Z_N\) two-form gauge field on \(Y\).  The homotopy cardinality of a
finite \(2\)-groupoid is
\[
  \sum_{[b]\in\pi_0}
  \frac{|\pi_2(\mathcal Z_N^2(Y),b)|}
       {|\pi_1(\mathcal Z_N^2(Y),b)|}.
\]
Here
\[
  \pi_0\simeq H^2(Y,\mathbb Z_N),\qquad
  \pi_1\simeq H^1(Y,\mathbb Z_N),\qquad
  \pi_2\simeq H^0(Y,\mathbb Z_N),
\]
so
\begin{equation}
  |\mathcal Z_N^2(Y)|
  =
  \frac{|H^0(Y,\mathbb Z_N)|\,|H^2(Y,\mathbb Z_N)|}
       {|H^1(Y,\mathbb Z_N)|}.
  \label{eq:finite-two-form-groupoid-homotopy-cardinality}
\end{equation}
Equivalently, if one sums over the unquotiented set of closed cochains, the
same cardinality is
\begin{equation}
  |\mathcal Z_N^2(Y)|
  =
  \frac{|C^0(Y,\mathbb Z_N)|\,|Z^2(Y,\mathbb Z_N)|}
       {|C^1(Y,\mathbb Z_N)|}.
  \label{eq:finite-two-form-groupoid-cochain-cardinality}
\end{equation}
This is exactly the normalization used in
\eqref{eq:triangulated-one-form-condensation-defect}: the summand
\(\mathsf U_b(Y)\) is written for every closed cochain \(b\), while the factor
\(|C^0|/|C^1|\) divides by one-form gauge redundancy and restores the
zero-form gauge-for-gauge volume.  Thus the normalization is not an optional
cellular convention; it is the finite path-integral measure for the
two-form gauge field.

\begin{theorem}[Finite higher gauging and condensation fusion]
\label{thm:finite-higher-gauging-condensation-defects}
Under the hypotheses of
Definition~\ref{def:triangulated-higher-gauging-defect}, the insertion
\(\mathfrak C(Y)\) is a topological codimension-one defect implementing
finite \(\mathbb Z_N\) one-form gauging on the hypersurface \(Y\).  The
fusion of two parallel copies is
\begin{equation}
  \mathfrak C(Y)\circ\mathfrak C(Y)
  =
  Z^{(2)}_{\mathbb Z_N}(Y)\,\mathfrak C(Y),
  \qquad
  Z^{(2)}_{\mathbb Z_N}(Y)
  =
  \frac{|C^0(Y,\mathbb Z_N)|\,|Z^2(Y,\mathbb Z_N)|}
       {|C^1(Y,\mathbb Z_N)|}.
  \label{eq:finite-higher-gauging-fusion}
\end{equation}
Equivalently,
\begin{equation}
  Z^{(2)}_{\mathbb Z_N}(Y)
  =
  \frac{|H^0(Y,\mathbb Z_N)|\,|H^2(Y,\mathbb Z_N)|}
       {|H^1(Y,\mathbb Z_N)|},
  \label{eq:finite-higher-gauging-topological-factor}
\end{equation}
so the coefficient is the closed-three-manifold partition value of the
finite three-dimensional \(\mathbb Z_N\) two-form gauge theory in this
normalization.  With corners or boundaries the same state sum assigns the
corresponding Hilbert spaces and operators; hence the fusion coefficient of
condensation defects is, in general, a three-dimensional topological theory
rather than an integer.

If the half-space gauged theory is identified with \(\mathcal Q\) after all
discrete theta terms and background boundary conditions are fixed, the
half-space gauging interface is a duality defect of \(\mathcal Q\).  For
\(N>1\) it is noninvertible whenever the associated two-form gauge theory has
non-one-dimensional state spaces, for example on a connected closed spatial
surface where the flux label in \(H^2(\Sigma,\mathbb Z_N)\) survives.
\end{theorem}

\begin{proof}
Topological invariance of \(\mathfrak C(Y)\) has two logically distinct
inputs.  The QFT input is that the surface networks \(U_n(\Sigma)\), their
junctions, and their Pachner-move identifications have already been
constructed as anomaly-free topological operators.  The finite gauge-theory
input is the groupoid measure
\eqref{eq:finite-two-form-groupoid-cochain-cardinality}.  The cochain
condition \(\delta b=0\) is exactly the junction conservation law for the
Poincare-dual network, so a closed \(b\) produces a network without residual
endpoints.  A change \(b\mapsto b+\delta\lambda\) inserts the boundary of a
one-cochain of symmetry surfaces; by the assumed topological junction
relations this is a gauge-equivalent representative of the same condensed
configuration.  A change \(\lambda\mapsto\lambda+\delta\xi\) is the
gauge-for-gauge redundancy of that boundary filling.  Therefore the sum in
\eqref{eq:triangulated-one-form-condensation-defect} is precisely the
unquotiented expression for the finite path integral over
\(\mathcal Z_N^2(Y)\).  In a collar
\((- \epsilon,\epsilon)\times Y\), replacing the fixed one-form background by
this groupoid sum is the finite gauging operation on the hypersurface.  This
proves the gauging interpretation after the QFT surface-operator input has
been supplied.

Place two copies of the defect on nearby parallel copies of \(Y\).  Their
product is
\[
  \left(\frac{|C^0(Y,\mathbb Z_N)|}{|C^1(Y,\mathbb Z_N)|}\right)^2
  \sum_{b,b'\in Z^2(Y,\mathbb Z_N)}
  \mathsf U_b(Y)\,\mathsf U_{b'}(Y).
\]
Using the junction isomorphism \(U_nU_m\simeq U_{n+m}\), change variables to
\[
  s=b+b',\qquad t=b .
\]
For each closed \(s\), the variable \(t\) runs freely over
\(Z^2(Y,\mathbb Z_N)\).  The product therefore becomes
\[
  \frac{|C^0|\,|Z^2|}{|C^1|}
  \left(
    \frac{|C^0|}{|C^1|}
    \sum_{s\in Z^2(Y,\mathbb Z_N)}\mathsf U_s(Y)
  \right),
\]
where all cochain groups are on \(Y\).  The second parenthesis is
\(\mathfrak C(Y)\), and the first factor is
\eqref{eq:finite-higher-gauging-fusion}.

It remains to see that this factor is topological.  From the short exact
sequences
\[
  0\to Z^1\to C^1\to B^2\to0,\qquad
  0\to B^1\to Z^1\to H^1\to0,\qquad
  0\to H^0\to C^0\to B^1\to0,
\]
and
\[
  0\to B^2\to Z^2\to H^2\to0,
\]
finite cardinalities multiply, giving
\[
  \frac{|C^0|\,|Z^2|}{|C^1|}
  =
  \frac{|H^0|\,|H^2|}{|H^1|}.
\]
This proves \eqref{eq:finite-higher-gauging-topological-factor}.  The
extension to manifolds with boundary is the same finite gauge-theory state
sum with boundary cochains left unsummed or summed according to the chosen
boundary condition; it assigns vector spaces and linear maps, with closed
three-manifold numbers as the special case.  If half-space gauging lands back
on \(\mathcal Q\), the interface has the same theory on both sides and is
therefore an internal
duality defect.  Its product with the opposite interface contains the
nontrivial two-form gauge-theory summation just computed, so it is invertible
only when that finite topological theory is the trivial invertible one.
\end{proof}

\paragraph{Relative state space of the finite factor.}
The last sentence of the theorem has concrete content.  Let \(\Sigma\) be a
closed oriented triangulated surface obtained by cutting the three-dimensional
defect worldvolume.  A boundary value of the finite two-form gauge field is a
closed two-cochain
\[
  z\in Z^2(\Sigma,\mathbb Z_N).
\]
Since \(\dim \Sigma=2\), this is the same set as \(C^2(\Sigma,\mathbb Z_N)\).
The boundary one-form gauge transformations identify
\[
  z\sim z+\delta\lambda,\qquad
  \lambda\in C^1(\Sigma,\mathbb Z_N).
\]
Thus the flux sectors on the cut surface are
\begin{equation}
  [z]\in H^2(\Sigma,\mathbb Z_N),
  \qquad
  \mathcal H^{(2)}_N(\Sigma)
  =
  \mathbb C\bigl[H^2(\Sigma,\mathbb Z_N)\bigr],
  \label{eq:finite-two-form-boundary-state-space}
\end{equation}
with the automorphism groups \(H^0\) and \(H^1\) entering the gluing pairing
through the same homotopy cardinality weights as in
\eqref{eq:finite-two-form-groupoid-homotopy-cardinality}.  More explicitly, a
cobordism \(W:\Sigma_0\to\Sigma_1\) gives a matrix whose
\(([z_1],[z_0])\)-entry is the homotopy cardinality of the finite groupoid of
closed two-cochains \(b\in Z^2(W,\mathbb Z_N)\) restricting to the prescribed
boundary classes, with one-cochain gauge transformations and zero-cochain
gauge-for-gauge transformations fixed on the boundary.  For the cylinder
\(\Sigma\times[0,1]\), restriction to the two ends forces
\([z_0]=[z_1]\), so the operator is diagonal on the basis
\eqref{eq:finite-two-form-boundary-state-space} up to the conventional
automorphism scalar fixed by the gluing normalization.

This relative description is what the closed-manifold number
\(Z^{(2)}_{\mathbb Z_N}(Y)\) becomes when a cut is present.  On a connected
closed oriented surface,
\[
  H^2(\Sigma,\mathbb Z_N)\simeq \mathbb Z_N,
\]
so the finite factor has an \(N\)-dimensional flux-sector space for \(N>1\).
Consequently fusion with the reverse gauging interface contains actual finite
topological degrees of freedom.  It is an invertible scalar line only when the
relevant finite state spaces are all one-dimensional.

For Yang--Mills theories this construction becomes a QFT statement only after
the \(\mathbb Z_N\) one-form symmetry surfaces, their junctions, their
correlation topology, and their anomaly trivialization have been constructed
inside the theory or by a regulator such as the lattice.  Once that input is
available, the cochain calculation above is the local algebraic mechanism
behind higher gauging and the resulting noninvertible condensation defects.

\section{Chiral Defects In Abelian Gauge Theory}
\label{sec:chiral-defects-abelian-gauge-theory}

We now give the explicit abelian construction.  Work on an oriented spin
four-manifold and take compact unit-charge massless QED with gauge field
\(A\), curvature \(F=\dd A\), and flux normalization
\[
  \int_\Sigma \frac{F}{2\pi}\in\mathbb Z
\]
for every closed two-cycle \(\Sigma\).  Let \(j_5\) be the axial current in the
normalization of Chapter~\ref{ch:volume-ii-20}: the axial rotation
by angle \(\alpha\) acts on the Dirac field by
\[
  \psi\mapsto \exp(\ii\alpha\gamma_5/2)\psi ,
\]
and, in the massless theory,
\begin{equation}
  \dd\star j_5=\frac{1}{4\pi^2}F\wedge F .
  \label{eq:qed-form-axial-anomaly-for-chiral-defect}
\end{equation}
This is the differential-form version of
\eqref{eq:abelian-axial-anomaly} with \(e=q=1\).  The magnetic one-form
symmetry has current
\[
  j_m=\frac{1}{2\pi}\star F,
  \qquad
  \dd\star j_m=\frac{1}{2\pi}\dd F=0 .
\]
It is part of the QFT definition only when no dynamical monopole operator has
been included; an inserted or dynamical monopole is an endpoint for the charged
't Hooft line and explicitly breaks this symmetry.

For a closed oriented three-manifold \(M\) supporting a defect, the naive axial
wall
\[
  U_\alpha(M)
  =
  \exp\left(\frac{\ii\alpha}{2}\int_M\star j_5\right)
\]
is not topological.  If \(M\) is moved across a four-chain \(W\),
\eqref{eq:qed-form-axial-anomaly-for-chiral-defect} changes it by
\[
  \exp\left(\frac{\ii\alpha}{8\pi^2}\int_W F\wedge F\right).
\]
It is also not a well-defined compact-gauge-field operator after replacing
\(\star j_5\) by the formally conserved Chern--Simons-improved current,
because the required fractional Chern--Simons level is not invariant under
large gauge transformations on a general compact \(M\).

\subsection{The Fractional Hall Dressing}

Let \(N>0\), let \(p\in\mathbb Z\) with \(\gcd(p,N)=1\), and put
\[
  \alpha=\frac{2\pi p}{N}.
\]
Let \(\mathcal A_{N,p}\) denote the minimal abelian three-dimensional TQFT with
a \(\mathbb Z_N\) one-form symmetry whose generating line \(\ell\) has spins
\[
  h(\ell^s)=\frac{p s^2}{2N}\quad \bmod 1 .
\]
The spin refinement of \(\mathcal A_{N,p}\) is included when the pair
\((N,p)\) requires it; the spin structure on \(M\) is inherited from the
ambient spin four-manifold.  Equivalently, coupling this one-form symmetry to
a \(\mathbb Z_N\)-valued two-cocycle \(B\) has four-dimensional inflow
\[
  \exp\left(
  -\frac{2\pi\ii p}{2N}\int_Y P(B)
  \right),
\]
on a four-manifold \(Y\) extending the background; \(P\) is the
Pontryagin-square cohomology operation.  When the integral curvature class of
\(A|_M\) is reduced modulo \(N\), it defines
\[
  B_A=\left[\frac{F}{2\pi}\right]\bmod N
  \in H^2(M,\mathbb Z_N).
\]
The dressed chiral defect is
\begin{equation}
  D_{N,p}(M)
  =
  \exp\left(\frac{\pi\ii p}{N}\int_M\star j_5\right)
  Z_{\mathcal A_{N,p}}(M;B_A).
  \label{eq:qed-noninvertible-chiral-defect-general}
\end{equation}
For \(p=1\), the minimal theory is the \(U(1)_N\) Chern--Simons theory, and
\eqref{eq:qed-noninvertible-chiral-defect-general} becomes the more concrete
path integral
\begin{equation}
  D_{N,1}(M)
  =
  \int\mathcal D a\,
  \exp\left[
  \frac{\pi\ii}{N}\int_M\star j_5
  +
  \ii\int_M
  \left(
    \frac{N}{4\pi}a\wedge\dd a
    +\frac{1}{2\pi}a\wedge\dd A
  \right)
  \right],
  \label{eq:qed-noninvertible-chiral-defect-u1n}
\end{equation}
where \(a\) is a compact \(U(1)\) gauge field living only on the defect
worldvolume.  The field \(a\) is not a new bulk particle; it is part of the
operator insertion.

\paragraph{Topological cancellation for \(D_{N,1}\).}
In the massless compact QED path integral without dynamical monopoles, the
operator \eqref{eq:qed-noninvertible-chiral-defect-u1n} is invariant under
smooth deformations of \(M\) that avoid charged insertions.

Let \(M_0\) and \(M_1\) be two homologous placements of the defect and let
\(W\) be the oriented four-chain swept out by the deformation.  The axial
factor in \eqref{eq:qed-noninvertible-chiral-defect-u1n} changes by
  \[
  \exp\left(\frac{\pi\ii}{N}\int_W\dd\star j_5\right)
  =
  \exp\left(\frac{\ii}{4\pi N}\int_W F\wedge F\right).
\]
The \(U(1)_N\) defect theory is gauge invariant as a compact
Chern--Simons theory.  To compute its local dependence on the ambient field
\(A\), one may integrate out \(a\) at the level of the response action:
the equation of motion is \(N\dd a+\dd A=0\), so locally
\[
  \frac{N}{4\pi}a\wedge\dd a+\frac{1}{2\pi}a\wedge\dd A
  \quad\leadsto\quad
  -\frac{1}{4\pi N}A\wedge\dd A .
\]
Therefore the change of the Hall response under the same deformation is
\[
  \exp\left(
  -\frac{\ii}{4\pi N}\int_W F\wedge F
  \right),
\]
which cancels the axial-anomaly phase.  The argument uses only the local
response to a deformation; the compact \(a\)-path integral in
\eqref{eq:qed-noninvertible-chiral-defect-u1n} is what makes the operator
globally well-defined under large gauge transformations.

For general \(p/N\), the same cancellation is expressed by the one-form anomaly
of \(\mathcal A_{N,p}\).  The Pontryagin-square response of
\(\mathcal A_{N,p}(B_A)\) is the compact, gauge-invariant replacement for the
fractional Chern--Simons counterterm whose differential-form shadow is
\[
  -\frac{\ii p}{4\pi N}\int A\wedge\dd A .
\]
Its deformation phase cancels the axial phase from
\eqref{eq:qed-form-axial-anomaly-for-chiral-defect}.  Thus
\eqref{eq:qed-noninvertible-chiral-defect-general} is the compact
topological defect associated with the rational axial angle \(2\pi p/N\).

\subsection{Half-Space Gauging Proof}

There is a second construction that explains why the defect is tied to the
magnetic one-form symmetry.  Let \(X_+=\{x\ge0\}\) and let \(M=\partial X_+\).
To gauge the \(\mathbb Z_N\) subgroup of the magnetic \(U(1)^{(1)}\) symmetry
in \(X_+\), introduce a compact two-form gauge field \(b\) and a compact
one-form gauge field \(c\) on \(X_+\).  For \(k\in\mathbb Z\) satisfying
\[
  pk\equiv1\pmod N,
\]
add
\begin{equation}
  S_{\mathrm{mag},N,k}
  =
  \ii\int_{X_+}
  \left(
    \frac{1}{2\pi}b\wedge F
    +\frac{N}{2\pi}b\wedge\dd c
    +\frac{Nk}{4\pi}b\wedge b
  \right).
  \label{eq:qed-half-space-magnetic-gauging}
\end{equation}
The \(c\)-equation imposes the flatness condition that makes \(b\) a
\(\mathbb Z_N\) two-form gauge field after the finite gauge quotient.  The
last term is the discrete-torsion choice.  In the smooth sector, completing
the square gives the theta-angle shift
\[
  \Delta S_\theta
  =
  -\frac{\ii p}{4\pi N}\int_{X_+}F\wedge F,
\]
with the integral quantization fixed by the compact \(b,c\) path integral.
The massless axial change of variables with
\[
  \alpha(x)=\frac{2\pi p}{N}\Theta(x)
\]
contributes the inverse phase and a boundary insertion
\[
  \exp\left(\frac{\pi\ii p}{N}\int_M\star j_5\right).
\]
After imposing the topological boundary condition \(b|_M=0\), the remaining
boundary theory produced by the \(b,c\) path integral is precisely
\(\mathcal A_{N,p}(B_A)\).  The interface between the ungauged side and the
half-space-gauged side is therefore
\eqref{eq:qed-noninvertible-chiral-defect-general}.  Since the operation is
gauging on a half-space followed by a change of variables, the interface is
topological as long as the magnetic one-form symmetry is unbroken.

\subsection{Fusion And Action On Operators}

The defect acts on local fermion operators by the axial rotation
\[
  \psi\mapsto \exp(\pi\ii p\gamma_5/N)\psi .
\]
Wilson lines are not charged under the magnetic one-form symmetry, so the
half-space gauging part leaves them unchanged.  A minimal 't Hooft line
\(H(\gamma)\), in the sense of
Definition~\ref{def:renormalized-thooft-line}, is charged.  If
\(\gamma=\partial\Sigma\) while the defect is swept past the line, gauge
invariance attaches the surface factor
\begin{equation}
  H(\gamma)\longmapsto
  H(\gamma)\,
  \exp\left(\frac{\ii p}{N}\int_\Sigma F\right),
  \label{eq:qed-chiral-defect-action-on-thooft-line}
\end{equation}
with \(\Sigma\) stretching between the line and the displaced defect.  This is
the operational reason the defect is not an ordinary axial symmetry operator:
its action on extended operators remembers the magnetic one-form symmetry.

The orientation-reverse fusion already shows noninvertibility.  For \(p=1\),
\begin{equation}
  D_{N,1}(M)D_{N,1}^\dagger(M)
  =
  \int\mathcal D a\,\mathcal D\bar a\,
  \exp\left[
  \ii\int_M
  \left(
    \frac{N}{4\pi}a\wedge\dd a
    -\frac{N}{4\pi}\bar a\wedge\dd\bar a
    +\frac{1}{2\pi}(a-\bar a)\wedge\dd A
  \right)
  \right].
  \label{eq:qed-dn1-orientation-reverse-fusion}
\end{equation}
The axial factors cancel, but the Hall theories do not cancel to the trivial
three-dimensional theory.  Equation
\eqref{eq:qed-dn1-orientation-reverse-fusion} is the magnetic
\(\mathbb Z_N\) condensation operator.  In the finite cochain language used
above, its general \(p\) analogue is a sum over \(\mathbb Z_N\)-valued
two-cocycles on \(M\), weighted by the Pontryagin-square phase and by the
corresponding magnetic one-form symmetry surface network.  Thus the fusion
coefficient is a three-dimensional TQFT or finite state sum, not a number.

\begin{controlledapproximation}[QED noninvertible chiral defects]
\label{ca:qed-noninvertible-chiral-defects}
In compact unit-charge massless QED on oriented spin four-manifolds, with no
dynamical monopoles and with the magnetic one-form symmetry retained in the
QFT definition, every rational axial angle \(2\pi p/N\) with
\(\gcd(p,N)=1\) defines the topological defect
\(D_{N,p}\) in \eqref{eq:qed-noninvertible-chiral-defect-general}.  The defect
acts on local fermions by the corresponding axial rotation, acts on 't Hooft
lines by \eqref{eq:qed-chiral-defect-action-on-thooft-line}, and is
noninvertible because
\(D_{N,p}D_{N,p}^{\dagger}\) is a magnetic condensation defect rather than the
invisible defect.
\end{controlledapproximation}

This statement is the compact path-integral construction just given.  A fully
axiomatic theorem would still require constructing compact
QED, its charged line sectors, and the defect Hilbert spaces in a
nonperturbative reconstruction framework.  If dynamical monopoles are added,
the magnetic one-form symmetry has endpoints and the half-space gauging step
is no longer topological; the defect then fails to define an exact symmetry.

Let a more general four-dimensional abelian gauge theory contain massless
fermions with a classical chiral transformation \(R_\alpha\).  Suppose the ABJ
anomaly changes the action by
\[
  \Delta S_\alpha
  =
  \frac{i\alpha q}{8\pi^2}\int_M F\wedge F ,
\]
where \(F\) is the \(U(1)\) field strength and \(q\in\mathbb Z\) is the anomaly
coefficient in the chosen charge normalization.  The codimension-one defect
implementing \(R_\alpha\) is topological only if this phase is cancelled or
trivialized as the defect crosses insertions.  In the absence of dynamical
monopoles, the theory also has a magnetic one-form symmetry, and the anomaly
phase can be represented by coupling the defect worldvolume to a
three-dimensional topological theory whose anyons attach to magnetic
background flux sectors.  The resulting dressed defect is generally
noninvertible: fusing it with its orientation reverse leaves a sum or a
three-dimensional topological factor rather than the invisible defect alone.

\section{Higher Fusion Categories And Truncations}
\label{sec:higher-fusion-categories-truncations}

The categorical level depends on which defects are retained.  In a
four-dimensional QFT, the full wall-junction system includes walls, junction
surfaces, junction lines, and point junctions.  It is therefore at least a
three-categorical defect object.  A fusion two-category appears only after a
controlled truncation, for example by focusing on topological surfaces and
their line and point junctions, or by studying the three-dimensional theory
living on a wall.  A finite semisimple fusion two-category has objects,
one-morphisms, and two-morphisms, a monoidal product, duals, and finite
semisimple Hom categories.  Condensation is then formulated by separable
algebra objects and their module two-categories.

This distinction matters in four-dimensional generalized symmetry.  A
one-form symmetry generated by topological surfaces is not fully described by
a one-category of simple surfaces if its line junctions carry nontrivial
operator spaces.  The symmetry TQFT keeps this extra information because bulk
boundary-changing defects, their junctions, and their endpoint operators live
one categorical level higher than the ordinary group law.  When a chapter later
uses only a fusion ring of surfaces, the omission is a truncation and not a
definition of the full symmetry object.

\section{Example: Mixed Zero- And One-Form Data}

Consider a four-dimensional gauge theory with gauge group \(SU(N)\) and a
charge conjugation symmetry acting on the center one-form symmetry
\[
  A=\mathbb Z_N
\]
by inversion.  A background for the ordinary symmetry changes the local
system in which the \(A\)-valued two-form background lives.  If a theta-angle
or discrete theta term changes under charge conjugation, the anomaly is a
five-dimensional functional of the combined background.  The line-operator
lattice then transforms as a module over the two-group rather than as a
separate representation of each symmetry.

\section*{Comparison Summary}

The symmetry-TQFT description is not a replacement for constructing the QFT.
It records:
\[
\begin{gathered}
  \text{allowed background fields},\qquad
  \text{anomaly lines},\\
  \text{gauging and condensation operations},\qquad
  \text{topological defect fusion}.
\end{gathered}
\]
Local correlators, Hilbert space, stress tensor, and dynamical phase
structure must still be supplied by the QFT itself.

\begin{openproblem}[Higher-group examples]
\label{op:higher-group-examples}
Construct interacting continuum QFT examples whose higher-group backgrounds,
symmetry TQFT, defect fusion, anomaly inflow, and local-operator action are
all derived from a single regulator or reconstruction framework.
\end{openproblem}
