\chapter{Duality Defects, Gauging, And Orbifold Data}
\label{ch:global-duality-defects-gauging-orbifold-data}

\ConventionsMixed

Gauging a finite symmetry is a finite path integral over background fields.
A duality defect is an interface obtained from such a gauging operation,
possibly followed by an equivalence between the gauged theory and another
theory.  The definitions in this chapter are formulated in terms of
background-field groupoids, topological symmetry defects, and line-charge
lattices.  This keeps separate three pieces of data which are often
compressed in the literature: the finite sum over bundles, the anomaly or
counterterm choice, and the identification of the resulting theory with a
theory already under study.

\section{Finite Symmetry Backgrounds}

Let \(G\) be a finite internal symmetry of a \(D\)-dimensional QFT
\(\mathcal Q\).  A background \(G\)-field on a spacetime \(M\) is a principal
\(G\)-bundle
\[
  P\to M.
\]
On a good cover \(\{U_i\}\) it may be represented by locally constant
transition functions
\[
  g_{ij}:U_i\cap U_j\longrightarrow G
\]
obeying
\[
  g_{ij}g_{jk}=g_{ik}
  \quad\text{on }U_i\cap U_j\cap U_k.
\]
Two representatives related by functions \(u_i:U_i\to G\),
\[
  g_{ij}\mapsto u_i g_{ij}u_j^{-1},
\]
define isomorphic bundles.  Equivalently, after choosing a triangulation, a
background is a flat \(G\)-valued one-cocycle on edges, modulo vertex gauge
transformations.

The QFT with background \(P\) has partition function
\[
  Z_{\mathcal Q}[M;P],
\]
and correlation functions with insertions carry the same background datum.
The collection of bundles and their isomorphisms is the finite groupoid
\[
  \operatorname{Bun}_G(M).
\]
For a representative \(P\), the automorphism group
\[
  \operatorname{Aut}(P)
\]
is the group of gauge transformations preserving \(P\).  On connected \(M\)
and for the trivial bundle it is \(G\), while for a general flat bundle it is
the centralizer of the holonomy representation.

\begin{definition}[Finite gauging groupoid sum]
\label{def:finite-gauging-groupoid-sum}
Assume the \(G\)-symmetry is anomaly-free after all local counterterms have
been fixed.  The gauged partition function is the groupoid cardinality sum
\begin{equation}
  Z_{\mathcal Q/G}(M)
  =
  \sum_{[P]\in\pi_0\operatorname{Bun}_G(M)}
  \frac{Z_{\mathcal Q}[M;P]}{|\operatorname{Aut}(P)|}.
  \label{eq:finite-gauging-groupoid-sum}
\end{equation}
Different choices of an invertible finite gauge theory multiplying
\eqref{eq:finite-gauging-groupoid-sum} are different normalization
conventions.  Once a convention is fixed, the same convention must be used for
closed manifolds, manifolds with boundary, interfaces, and operator
insertions.
\end{definition}

The groupoid factor is part of the finite path integral.  It is the finite
analogue of dividing by the gauge group volume in a continuous gauge theory.

\begin{remark}[Disjoint-union multiplicativity]
\label{rem:finite-gauging-disjoint-union}
With the groupoid convention \eqref{eq:finite-gauging-groupoid-sum}, finite
gauging is multiplicative under disjoint union:
\[
  Z_{\mathcal Q/G}(M_1\sqcup M_2)
  =
  Z_{\mathcal Q/G}(M_1)Z_{\mathcal Q/G}(M_2),
\]
provided the original background partition functions are multiplicative.

A principal \(G\)-bundle on \(M_1\sqcup M_2\) is a pair
\((P_1,P_2)\) with \(P_i\to M_i\).  Its automorphism group is the product
\[
  \operatorname{Aut}(P_1,P_2)
  =
  \operatorname{Aut}(P_1)\times\operatorname{Aut}(P_2).
\]
Thus the groupoid sum factorizes:
\[
\begin{aligned}
  Z_{\mathcal Q/G}(M_1\sqcup M_2)
  &=
  \sum_{[P_1],[P_2]}
  \frac{
    Z_{\mathcal Q}[M_1;P_1]Z_{\mathcal Q}[M_2;P_2]
  }{
    |\operatorname{Aut}(P_1)|\,|\operatorname{Aut}(P_2)|
  }  \\
  &=
  Z_{\mathcal Q/G}(M_1)Z_{\mathcal Q/G}(M_2).
\end{aligned}
\]
\end{remark}

\section{Dijkgraaf--Witten Twists And Anomaly Data}

A class
\[
  \omega\in H^D(BG,U(1))
\]
defines an invertible topological action for a \(G\)-bundle \(P\to M\).  If
\(\Omega\) is a cocycle representative, the exponentiated action is
\[
  \exp(\ii S_\omega(P))
  =
  \exp\left(2\pi\ii\int_M f_P^*\Omega\right),
\]
where \(f_P:M\to BG\) classifies \(P\).  On a triangulated manifold this is
the evaluation of the finite group cocycle on the flat edge labels of \(P\).
The twisted gauging is
\begin{equation}
  Z_{\mathcal Q/G,\omega}(M)
  =
  \sum_{[P]}
  \frac{Z_{\mathcal Q}[M;P]\exp(\ii S_\omega(P))}
       {|\operatorname{Aut}(P)|}.
  \label{eq:dw-twisted-gauging}
\end{equation}

If \(\mathcal Q\) has a \(G\)-anomaly
\[
  \alpha\in H^{D+1}(BG,U(1)),
\]
then \(Z_{\mathcal Q}[M;P]\) is naturally a vector in a one-dimensional
anomaly line rather than an ordinary complex number.  Gauging requires a
trivialization of this anomaly line over the groupoid of \(G\)-bundles, or a
\((D+1)\)-dimensional invertible bulk whose boundary is \(\mathcal Q\).  A
Dijkgraaf--Witten twist in degree \(D\) changes the gauged theory; it cancels
an anomaly only through its coboundary when a cochain-level trivialization of
the degree \(D+1\) anomaly has been chosen.

\section{The Gauging Interface}

There is a codimension-one interface
\[
  \mathcal N_G:\mathcal Q\longrightarrow \mathcal Q/G
\]
obtained by gauging \(G\) on one side of the interface.  More concretely, if a
spacetime is cut along a hypersurface \(Y\), then on the gauged side one sums
over \(G\)-bundles whose restriction to \(Y\) is allowed to vary, while on the
ungauged side the boundary value is interpreted as a background for
\(\mathcal Q\).  The reverse interface is denoted
\[
  \mathcal N_G^\dagger:\mathcal Q/G\longrightarrow\mathcal Q .
\]

For every \(g\in G\) let \(D_g\) be the invertible symmetry defect of
\(\mathcal Q\).  It is supported on \(Y\) and imposes \(G\)-holonomy \(g\)
across a small transverse interval.  The orientation convention is
\[
  D_g\otimes D_h\simeq D_{gh}.
\]

\begin{proposition}[Slab derivation of the gauging-interface fusion]
\label{prop:gauging-interface-slab-fusion}
With compatible groupoid normalizations for the two interfaces,
\begin{equation}
  \mathcal N_G^\dagger\circ\mathcal N_G
  \simeq
  \bigoplus_{g\in G}D_g
  \label{eq:gauging-interface-slab-fusion}
\end{equation}
as defects of \(\mathcal Q\).
\end{proposition}

\begin{proof}
Fuse \(\mathcal N_G\) with \(\mathcal N_G^\dagger\) and separate the two
interfaces by a thin slab \(Y\times[0,\epsilon]\).  Inside the slab the
\(G\)-field is flat.  Choose trivializations on the two boundary components
of the slab.  A flat bundle on an interval with these endpoint
trivializations is classified by a single holonomy \(g\in G\).  In the
sector with holonomy \(g\), transporting an operator across the slab inserts
the symmetry defect \(D_g\) on \(Y\).  Summing over the finite gauge field in
the slab therefore produces the direct sum over all holonomies.

The only normalization issue is the volume of gauge transformations in the
slab that are identity on the two boundary components.  With the interface
normalization induced from Definition~\ref{def:finite-gauging-groupoid-sum},
these factors cancel between the two sides of the cut, leaving one copy of
each holonomy sector.  Hence the fused interface is
\(\oplus_{g\in G}D_g\).
\end{proof}

The opposite fusion
\[
  \mathcal N_G\circ\mathcal N_G^\dagger
\]
is a defect of the gauged theory.  It is the algebra object generated by the
quantum symmetry defects acting on twisted sectors.

\begin{definition}[Regular algebra defect of a finite symmetry]
\label{def:regular-algebra-defect}
The regular algebra object associated to a finite anomaly-free symmetry \(G\)
is
\[
  A_G=\bigoplus_{g\in G}D_g
\]
with multiplication induced by the defect junctions
\[
  D_g\otimes D_h\longrightarrow D_{gh}
\]
and unit \(\mathbf 1=D_e\hookrightarrow A_G\).  As an object in the fusion
category generated by the symmetry defects,
\begin{equation}
  A_G\otimes A_G\simeq |G|\,A_G .
  \label{eq:regular-algebra-square}
\end{equation}
The displayed square follows directly from \(D_g\otimes D_h\simeq D_{gh}\):
\[
  A_G\otimes A_G
  \simeq
  \bigoplus_{g,h\in G}D_{gh}.
\]
For each fixed \(k\in G\), the equation \(gh=k\) has exactly \(|G|\)
solutions, one for each choice of \(g\), with \(h=g^{-1}k\).  Therefore every
simple summand \(D_k\) appears with multiplicity \(|G|\), which gives
\(|G|A_G\).
\end{definition}

\section{Orbifold Local Operators In Two Dimensions}

For a two-dimensional QFT with finite internal symmetry \(G\), a local
operator of the gauged theory is described by a punctured disk carrying a flat
\(G\)-bundle.  Choose a small positively oriented circle around the insertion.
Its holonomy is an element
\[
  g\in G.
\]
Changing the trivialization at the base point conjugates \(g\), so the
topological sector is a conjugacy class \([g]\).  Let
\[
  \mathcal V_g
\]
be the local operator space in the \(g\)-twisted sector.  Gauge
transformations preserving the chosen holonomy form the centralizer
\[
  C_G(g)=\{h\in G\mid hg=gh\}.
\]
Thus the local operator space of the gauged theory is
\begin{equation}
  \mathcal V_{\mathcal Q/G}
  =
  \bigoplus_{[g]\subset G}
  \mathcal V_g^{C_G(g)}.
  \label{eq:orbifold-local-operator-space}
\end{equation}

In a two-dimensional orbifold by a finite group \(G\), an OPE coefficient
between twisted sectors can receive a contribution from a flat bundle on a
pair of pants only if, after choosing compatible base paths,
\begin{equation}
  g_1g_2g_3^{-1}=e.
  \label{eq:orbifold-pair-of-pants-monodromy}
\end{equation}
Equivalently, the outgoing holonomy is the product \(g_3=g_1g_2\), up to
simultaneous conjugation.
The fundamental group of a pair of pants has generators represented by the
three boundary circles with the relation
\[
  \gamma_1\gamma_2\gamma_3^{-1}=1
\]
for the convention in which \(\gamma_3\) is the outgoing boundary.  A flat
\(G\)-bundle is a homomorphism from this fundamental group to \(G\), modulo
conjugation.  Therefore the boundary holonomies
\[
  g_i=\operatorname{Hol}(\gamma_i)
\]
must satisfy \(g_1g_2g_3^{-1}=e\).  Conversely any triple satisfying this
relation defines a flat bundle on the pair of pants after choosing a
trivialization at the base point.  Changing the base trivialization conjugates
all three holonomies simultaneously.

Formula \eqref{eq:orbifold-local-operator-space} and
the pair-of-pants monodromy condition \eqref{eq:orbifold-pair-of-pants-monodromy} give the finite
topological part of the orbifold OPE.  The analytic OPE coefficients inside
each allowed twisted sector are still QFT data; they are not determined by
the group \(G\) alone.

\section{Duality Defects From Self-Dual Gauging}

A duality defect is a topological interface whose fusion with its reverse is
the regular algebra defect \(A_G\) rather than a single copy of the invisible
defect.  The construction is especially transparent when a gauged theory is
equivalent to the original theory.

\begin{proposition}[Self-dual gauging produces a noninvertible defect]
\label{prop:self-dual-gauging-noninvertible-defect}
Let \(H\) be finite.  Assume that \(H\) acts internally on a QFT
\(\mathcal Q\) and can be gauged without an anomaly.  Suppose there is an
equivalence
\[
  \Phi:\mathcal Q/H\longrightarrow \mathcal Q
\]
after all background choices, discrete torsion choices, and local
counterterms have been fixed.  Then the composite
\[
  D_H:=\Phi\circ\mathcal N_H
\]
is a topological defect of \(\mathcal Q\) satisfying
\begin{equation}
  D_H^\dagger\otimes D_H
  \simeq
  A_H
  =
  \bigoplus_{h\in H}D_h .
  \label{eq:self-dual-gauging-defect-fusion}
\end{equation}
If \(H\) is nontrivial and the \(D_h\) are distinct simple invertible defects,
then \(D_H\) is noninvertible and has quantum dimension \(\sqrt{|H|}\).
\end{proposition}

\begin{proof}
The equivalence \(\Phi\) is invertible as an interface, so
\[
  \Phi^\dagger\circ\Phi\simeq\mathbf 1_{\mathcal Q/H}.
\]
Therefore
\[
  D_H^\dagger\otimes D_H
  =
  \mathcal N_H^\dagger\circ\Phi^\dagger\circ\Phi\circ\mathcal N_H
  \simeq
  \mathcal N_H^\dagger\circ\mathcal N_H .
\]
Proposition~\ref{prop:gauging-interface-slab-fusion} identifies the last
defect with \(\oplus_{h\in H}D_h=A_H\), proving
\eqref{eq:self-dual-gauging-defect-fusion}.

If \(D_H\) were invertible, \(D_H^\dagger\otimes D_H\) would be a single copy
of the invisible defect.  For nontrivial \(H\) with distinct simple \(D_h\),
the right hand side of \eqref{eq:self-dual-gauging-defect-fusion} has
\(|H|\) simple summands, so \(D_H\) is not invertible.

Quantum dimension is multiplicative under fusion and additive under direct
sum.  Since each \(D_h\) is invertible, \(d(D_h)=1\).  Hence
\[
  d(D_H)^2
  =
  d(D_H^\dagger\otimes D_H)
  =
  \sum_{h\in H}d(D_h)
  =
  |H|.
\]
Positivity gives \(d(D_H)=\sqrt{|H|}\).
\end{proof}

\begin{example}[The Ising Kramers--Wannier defect]
\label{ex:ising-kw-as-self-dual-gauging}
At the critical Ising point the anomaly-free spin-flip symmetry is
\(H=\mathbb Z_2=\{e,\eta\}\).  The gauged theory is equivalent to the original
critical Ising CFT after the Kramers--Wannier self-duality identification is
chosen.  Proposition~\ref{prop:self-dual-gauging-noninvertible-defect} gives
\[
  \mathcal N^\dagger\otimes\mathcal N
  \simeq
  \mathbf 1\oplus\eta,
  \qquad
  d(\mathcal N)=\sqrt2,
\]
which is the categorical fusion rule used in
Chapter~\ref{ch:global-categorical-symmetry-defect-fusion}.  The statement is
conditional on the self-duality equivalence, while the fusion calculation
itself is the finite slab calculation above.
\end{example}

\section{Normal-Subgroup Gauging}

Let \(G\) be a finite internal symmetry of \(\mathcal Q\), and let
\[
  H\triangleleft G
\]
be a normal subgroup.  Assume the \(H\)-symmetry is anomaly-free after the
restriction of all anomaly and counterterm data has been fixed.  Gauging \(H\)
produces \(\mathcal Q/H\).

If the \(G\)-action on \(\mathcal Q\) is anomaly-compatible with the chosen
\(H\)-gauging, then \(\mathcal Q/H\) has a residual ordinary symmetry
\[
  G/H .
\]
For \(g\in G\), conjugation maps an \(H\)-bundle \(P\) to another \(H\)-bundle
\[
  P^g:=P\times_{H,c_g}H,
  \qquad c_g(h)=ghg^{-1}.
\]
This is well-defined because \(H\) is normal.  The original \(G\)-symmetry
identifies the \(H\)-background partition functions for \(P\) and \(P^g\),
up to the specified anomaly-compatible counterterm.  Therefore \(g\) acts on
the groupoid sum defining \(\mathcal Q/H\).  If \(g\in H\), this action is a
gauge transformation of the summed \(H\)-field and hence acts trivially on
gauge-invariant observables of the gauged theory.  The action therefore
factors through \(G/H\).  Compatibility of the anomaly and counterterm choices
ensures that the factorized action is associative rather than projective.

\begin{example}[Normal-subgroup interface and a finite test case]
\label{ex:normal-subgroup-gauging-defect}
Let \(G=S_3\) and let
\[
  H=A_3=\{e,r,r^2\}\triangleleft S_3.
\]
After anomaly-free \(H\)-gauging, the residual ordinary symmetry is
\[
  S_3/A_3\simeq\mathbb Z_2.
\]
The gauging interface obeys
\[
  D_h\otimes\mathcal N_H\simeq\mathcal N_H,
  \qquad h\in H,
\]
because an \(H\)-defect ending on the gauged side can be absorbed into the
summed \(H\)-bundle.  At a self-dual point with an equivalence
\(\Phi:\mathcal Q/H\to\mathcal Q\), the defect
\[
  D_H=\Phi\circ\mathcal N_H
\]
satisfies
\[
  D_H^\dagger\otimes D_H
  \simeq
  \mathbf 1\oplus D_r\oplus D_{r^2},
  \qquad
  d(D_H)=\sqrt3 .
\]
This example separates the finite algebra from the dynamical question of
whether a particular QFT possesses the required self-duality equivalence.
\end{example}

\section{Gauge-Theory Duality Walls}

In gauge theories, duality walls can act on electric-magnetic line-charge
lattices.  Let
\[
  \Gamma
\]
be a free abelian group of line charges equipped with an integral
antisymmetric Dirac pairing
\[
  \langle\ ,\ \rangle:\Gamma\times\Gamma\to\mathbb Z.
\]
A line-lattice duality wall is a topological wall whose induced map
\[
  M:\Gamma\longrightarrow\Gamma
\]
is an automorphism preserving this pairing:
\begin{equation}
  \langle Mx,My\rangle=\langle x,y\rangle
  \qquad x,y\in\Gamma.
  \label{eq:line-lattice-symplectic-condition}
\end{equation}
If the global form of the gauge group or a discrete theta angle selects a
proper set of genuine lines, the wall is a wall between the corresponding
theories only when \(M\) maps the genuine line data on one side to the genuine
line data on the other side.  Boundary degrees of freedom may be required to
cancel an anomaly mismatch.

\begin{example}[Four-dimensional abelian line-lattice duality walls]
\label{ex:abelian-line-lattice-duality-walls}
For four-dimensional \(U(1)\) gauge theory with electric and magnetic line
charges
\[
  \Gamma=\mathbb Z^2,
  \qquad
  x=(q_e,q_m),
\]
take the Dirac pairing
\[
  \langle x,y\rangle=q_e q'_m-q_m q'_e .
\]
The two line-lattice generators used here act on column vectors by
\begin{equation}
  S(q_e,q_m)=(q_m,-q_e),
  \qquad
  T(q_e,q_m)=(q_e+q_m,q_m).
  \label{eq:abelian-st-line-action}
\end{equation}
The \(S\)-wall exchanges Wilson and 't Hooft charges with the orientation
shown in \eqref{eq:abelian-st-line-action}.  The \(T\)-wall implements the
Witten-effect shift of electric charge by magnetic charge.
\end{example}

\begin{remark}[Pairing preservation for the \(S\) and \(T\) walls]
\label{rem:st-walls-preserve-dirac-pairing}
The transformations in Example~\ref{ex:abelian-line-lattice-duality-walls}
preserve the Dirac pairing on \(\Gamma\).

For \(Sx=(q_m,-q_e)\) and \(Sy=(q'_m,-q'_e)\),
\[
  \langle Sx,Sy\rangle
  =
  q_m(-q'_e)-(-q_e)q'_m
  =
  q_e q'_m-q_m q'_e
  =
  \langle x,y\rangle .
\]
For \(Tx=(q_e+q_m,q_m)\) and \(Ty=(q'_e+q'_m,q'_m)\),
\[
  \langle Tx,Ty\rangle
  =
  (q_e+q_m)q'_m-q_m(q'_e+q'_m)
  =
  q_e q'_m-q_m q'_e
  =
  \langle x,y\rangle .
\]
Thus both maps are automorphisms of the charge lattice with its
antisymmetric Dirac pairing.
\end{remark}

\begin{example}[Finite center-sensitive line lattice]
\label{ex:finite-center-line-lattice-duality-wall}
For a theory whose center-sensitive electric and magnetic charges reduce to
\[
  \mathcal C_N=\mathbb Z_N^{\rm e}\oplus\mathbb Z_N^{\rm m},
\]
the finite pairing is
\[
  B((e,m),(e',m'))=\frac{em'-me'}{N}\in\mathbb Q/\mathbb Z.
\]
The reductions of \(S\) and \(T\) modulo \(N\),
\[
  S(e,m)=(m,-e),\qquad T(e,m)=(e+m,m),
\]
preserve \(B\).  More generally
\[
  T^p(e,m)=(e+pm,m)
\]
preserves \(B\) and maps a magnetic condensate direction to the dyonic
direction used in oblique confinement:
\[
  (0,1)\mapsto(p,1).
\]
Thus the finite line-lattice action gives a precise algebraic bridge between
duality walls and the oblique-confinement diagnostic of
Chapter~\ref{ch:global-confinement-screening-oblique}.
\end{example}

\section{Logical Scope}

The finite statements above are theorem-level statements about groupoids of
finite bundles, topological symmetry defects, and line-charge lattices.  To
apply them to an interacting continuum QFT one must supply the QFT input:
existence of the symmetry defects as genuine topological extended operators,
reflection-positive state spaces with the required defect sectors, anomaly
trivializations, and any claimed equivalence \(\Phi:\mathcal Q/H\to\mathcal
Q\).  Once those data are supplied, the slab fusion
\(\mathcal N_H^\dagger\mathcal N_H=A_H\), the orbifold twisted-sector
decomposition, and the line-lattice pairing preservation are finite
consequences.

\begin{openproblem}[Continuum gauging interfaces]
\label{op:continuum-gauging-interfaces}
Construct gauging interfaces and duality defects in interacting continuum QFTs
from background-field sums, local operator sectors, defect junctions, anomaly
trivializations, and reflection-positive state spaces.  The construction
should prove topological invariance of the interface, the existence and
composition law of junction operators, and compatibility with the Hilbert
spaces obtained by cutting spacetime along the defect.
\end{openproblem}
