\chapter{Line, Surface, and Domain-Wall Operators}
\label{ch:global-line-surface-domain-wall-operators}

\ConventionsMixed

Extended operators are QFT data supported on submanifolds.  They are not
determined by local operators alone.  This chapter defines the basic objects
needed for later discussions of confinement, screening, duality, and anomaly
inflow.

\section{Defect Data}

Let \(X\) be a spacetime manifold and let \(Y\subset X\) be an embedded
submanifold.  A defect supported on \(Y\) is specified by:
\begin{enumerate}[label=(\roman*)]
\item a choice of fields or boundary conditions on the complement
  \(X\setminus Y\);
\item singular behavior or extra degrees of freedom along \(Y\);
\item a coupling between bulk and defect variables;
\item a renormalization prescription for correlation functions with the
  defect inserted;
\item an equivalence relation identifying defect representatives with the
  same correlation functionals.
\end{enumerate}
The codimension of \(Y\) determines whether the object is a local operator,
line, surface, domain wall, boundary condition, or higher defect.

\section{Wilson Lines}

Let \(G\) be a compact gauge group and \(R\) a finite-dimensional
representation.  For an oriented curve \(\gamma\), the Wilson line is the
parallel-transport functional
\[
  W_R(\gamma)=
  \operatorname{Tr}_R\,\operatorname{Pexp}
  \int_\gamma A .
\]
In a regulated quantum gauge theory this expression is a coordinate for a
renormalized line operator.  Its definition requires the global form of
\(G\), the representation \(R\), framing or endpoint data when needed, and a
renormalization scheme for the short-distance singularity along the line.
If \(\gamma\) is open, the endpoints must carry charged states or additional
operators so that the full insertion is gauge invariant.

For a closed curve, choose a base point \(p\in\gamma\) and write
\(\operatorname{Hol}_\gamma(A)\in G\) for parallel transport from \(p\) back
to \(p\).  Under a gauge transformation \(u:X\to G\),
\[
  \operatorname{Hol}_\gamma(A^u)
  =
  u(p)^{-1}\operatorname{Hol}_\gamma(A)u(p),
\]
so \(\operatorname{Tr}_R\operatorname{Hol}_\gamma(A)\) is gauge invariant.
For an interval \(\gamma=[p_-,p_+]\), parallel transport is a map
\[
  U_\gamma(A):R_{p_-}\longrightarrow R_{p_+},
  \qquad
  U_\gamma(A^u)=R(u(p_+))^{-1}U_\gamma(A)R(u(p_-)).
\]
Gauge invariance is then obtained after contracting the endpoint indices with
charged endpoint vectors or endpoint local operators in the dual
representations.  This endpoint bookkeeping is part of the line definition
because it determines whether the line is genuine, a boundary of a surface
operator, or an operator that changes the charged sector of the Hilbert
space.

\section{Magnetic and Dyonic Lines}

A magnetic line in four-dimensional gauge theory is defined by a prescribed
singularity of the gauge field near a curve.  Around a small sphere linking
the line, the magnetic flux is labelled by a cocharacter
\[
  m:U(1)\longrightarrow G
\]
modulo conjugation, or equivalently by an element of the magnetic cocharacter
lattice modulo the Weyl group.  A dyonic line carries both the magnetic label
\(m\) and an electric representation of the stabilizer of \(m\).  The allowed
electric and magnetic labels obey the Dirac pairing condition: the phase
obtained by moving one line around another must be trivial for mutually local
operators in the chosen theory.

The global form of the gauge group changes the allowed labels.  For example,
the electric lattice of \(SU(N)\) differs from that of \(PSU(N)\), and the
dual magnetic lattice changes accordingly.  Therefore the line-operator
spectrum is part of the definition of a gauge theory: it is fixed by the
global gauge group, the allowed bundles, and the chosen mutually local
electric-magnetic charge lattice.

\subsection{The Local Monopole Model}

Let \(m:U(1)\to T\subset G\) be a cocharacter and let
\[
  m_*:\ii\mathbb R\longrightarrow\mathfrak t
\]
be its differential.  We identify \(m_*\) with the element of
\(\mathfrak t\) obtained by evaluating the differential on \(\ii\).  On a
small oriented two-sphere linking the candidate line, use spherical
coordinates \((\vartheta,\varphi)\).  Cover the sphere by the northern and
southern patches \(U_N=\{\vartheta\ne\pi\}\) and
\(U_S=\{\vartheta\ne0\}\).  The embedded Dirac monopole connection is
\begin{equation}
  A_N=\frac{m_*}{2}(1-\cos\vartheta)\,\dd\varphi,
  \qquad
  A_S=-\frac{m_*}{2}(1+\cos\vartheta)\,\dd\varphi .
  \label{eq:thooft-local-dirac-patches}
\end{equation}
On the overlap,
\[
  A_N-A_S=m_*\,\dd\varphi .
\]
The transition function
\[
  g_{NS}(\varphi)=m(e^{\ii\varphi})
\]
satisfies
\[
  A_N=g_{NS}^{-1}A_Sg_{NS}+g_{NS}^{-1}\dd g_{NS}
\]
in the abelianized \(T\)-valued model, and in a nonabelian theory after the
choice of local reduction to \(T\).  Its curvature is
\begin{equation}
  F_m=\frac{m_*}{2}\sin\vartheta\,\dd\vartheta\wedge\dd\varphi,
  \qquad
  \int_{S^2}\frac{F_m}{2\pi}=m_* .
  \label{eq:thooft-local-flux}
\end{equation}
The equality in the second formula means equality in the cocharacter lattice
after pairing with characters of \(T\).

\begin{proposition}[Cocharacter criterion for the monopole singularity]
\label{prop:thooft-cocharacter-criterion}
The local model \eqref{eq:thooft-local-dirac-patches} defines a
\(G\)-connection on the punctured ball modulo the chosen singularity exactly
when \(m\in X_*(T)=\operatorname{Hom}(U(1),T)\).  For every character
\(\lambda\in X^*(T)\), a charge-\(\lambda\) Wilson probe transported across
the Dirac ray changes by the phase
\[
  \exp\!\bigl(2\pi\ii\langle\lambda,m\rangle\bigr),
\]
which is one by the definition of the character-cocharacter pairing.
\end{proposition}

\begin{proof}
The northern and southern gauge potentials define a single connection on the
punctured sphere precisely when the transition map on the equatorial overlap
is a single-valued smooth map \(S^1\to T\).  Such maps are exactly
cocharacters \(m:U(1)\to T\).  The relation
\(g_{NS}^{-1}\dd g_{NS}=m_*\,\dd\varphi\) gives the patching equation above.

Let a Wilson probe have \(T\)-weight \(\lambda\).  In the one-dimensional
weight space, \(g_{NS}(\varphi)\) acts by
\[
  \lambda(g_{NS}(\varphi))
  =
  \lambda(m(e^{\ii\varphi}))
  =
  e^{\ii\langle\lambda,m\rangle\varphi}.
\]
Taking \(\varphi\) once around the equator gives
\(\exp(2\pi\ii\langle\lambda,m\rangle)\).  Since
\(\lambda\circ m:U(1)\to U(1)\) is a group homomorphism, it has integer
degree \(\langle\lambda,m\rangle\).  Hence the phase is trivial.  Conversely,
a \(T\)-valued transition map with this single-valuedness property for all
characters is an element of \(X_*(T)\), because characters separate points of
the compact torus.
\end{proof}

Two cocharacters related by the Weyl group determine gauge-equivalent local
singularities: a gauge transformation whose value on the linking sphere
implements the Weyl representative conjugates one Cartan reduction to the
other.  The underlying topological class of the \(G\)-bundle on the linking
two-sphere records only the image of \(m\) in \(\pi_1(G)\).  The full
cocharacter class is finer data: it fixes an asymptotic reduction of the
connection and thereby determines the stabilizer group and electric labels of
a dyonic refinement.

\begin{definition}[Renormalized 't Hooft line]
\label{def:renormalized-thooft-line}
Let \(X\) be an oriented four-manifold, let \(\gamma\subset X\) be an oriented
embedded curve, and let \(G\) be a compact connected gauge group with maximal
torus \(T\) and Weyl group \(W\).  A bare magnetic line of magnetic charge
\[
  m\in\operatorname{Hom}(U(1),T)/W
\]
is defined by removing a tubular neighborhood \(N_\epsilon(\gamma)\) and
integrating over \(G\)-connections on \(X\setminus N_\epsilon(\gamma)\) whose
restriction to each small linking sphere \(S^2_p\) lies in the topological
class determined by \(m\).  In a local trivialization over
\(\gamma\times(S^2_p\setminus\{\text{Dirac ray}\})\), this means that the
connection is gauge-equivalent near the boundary to the embedded Dirac
monopole
\[
  F\sim \frac{m_*}{2}\sin\theta\,\dd\theta\wedge\dd\varphi ,
  \qquad
  \int_{S^2_p}\frac{F}{2\pi}=m_* ,
\]
where \(m_*\in\mathfrak t\) is the differential of the cocharacter in the
chosen normalization.  Gauge transformations are required to preserve the
cocharacter class on the boundary.  Let \(\mathcal A_m\) denote the resulting
space of regulated connections and let \(\mathcal G_m\) denote the
corresponding boundary-preserving gauge group.  The renormalized 't Hooft line
\(H_m(\gamma)\) is the limit
\[
  H_m(\gamma)
  =
  \lim_{\epsilon\to0}
  Z_{\rm line}(\epsilon)\,
  \int_{\mathcal A_m(X\setminus N_\epsilon(\gamma))/\mathcal G_m}
  \mathcal D A\,\exp(-S[A])
\]
as a correlation functional with the same additional insertions, after
multiplication by local line counterterms \(Z_{\rm line}(\epsilon)\).  The
definition exists only in a specified regulator or constructive framework in
which the limit and the allowed counterterms are part of the line-operator
data.
\end{definition}

\subsection{Dyonic Refinement and the Stabilizer}

The magnetic singularity \(m\) breaks the gauge transformations along the
line to the centralizer
\[
  G_m=Z_G(m(U(1))).
\]
Equivalently, \(G_m\) is the compact subgroup whose Lie algebra is the sum of
\(\mathfrak t\) and the root spaces with roots \(\alpha\) satisfying
\(\langle\alpha,m\rangle=0\).  The tangential component of the connection
along \(\gamma\), restricted to the boundary condition above, is a
\(G_m\)-connection up to terms fixed by the singular normal model.  A dyonic
Wilson--'t Hooft line is therefore specified by
\[
  (m,\rho),
  \qquad
  m\in X_*(T)/W,\quad
  \rho\in\operatorname{Rep}(G_m),
\]
with the convention that \(\rho\) is replaced by its Weyl-conjugate
representation when \(m\) is Weyl-conjugated.  The insertion is the magnetic
boundary condition together with the Wilson parallel transport in \(\rho\)
for the residual \(G_m\)-connection along \(\gamma\), followed by the same
line-counterterm renormalization appearing in
Definition~\ref{def:renormalized-thooft-line}.  If \(\gamma\) has endpoints,
the endpoint states must be modules for the corresponding residual
stabilizer groups.

For center-sensitive questions it is often enough to abelianize the label to
\[
  (\lambda,m)\in
  \bigl(\Lambda_{\rm wt}/\Lambda_{\rm rt}\bigr)
  \oplus
  \bigl(\Lambda_{\rm cwt}/\Lambda_{\rm crt}\bigr),
\]
as in Chapter~\ref{ch:global-forms-higher-form-symmetry}.  This quotient
records the charge detected by electric and magnetic one-form symmetries.  The
full operator also contains the stabilizer representation, defect
counterterms, and the choice of regularization.

\begin{proposition}[Linking phase and mutual locality]
\label{prop:thooft-linking-phase}
Let \(L_{\lambda,m}\) and \(L_{\lambda',m'}\) be two abelianized dyonic
labels in the universal center-sensitive lattice.  In a configuration where
the first line links once with the second, the finite topological phase
obtained by moving the first line around the second is
\[
  \exp\!\left[
    2\pi\ii\bigl(\lambda(m')-\lambda'(m)\bigr)
  \right].
\]
Consequently two line labels can belong to the same mutually local genuine
line spectrum only if
\begin{equation}
  \lambda(m')-\lambda'(m)\in\mathbb Z .
  \label{eq:thooft-mutual-locality-condition}
\end{equation}
\end{proposition}

\begin{proof}
Choose disjoint tubular neighborhoods of the two curves.  Near
\(L_{\lambda',m'}\), the first line's Wilson factor of weight \(\lambda\)
sees the transition function \(m'(e^{\ii\varphi})\) when its path crosses the
Dirac ray of the second magnetic singularity.  By the computation in
Proposition~\ref{prop:thooft-cocharacter-criterion}, the phase is
\(\exp(2\pi\ii\lambda(m'))\).  Interchanging the roles of the two lines gives
\(\exp(2\pi\ii\lambda'(m))\).  The orientation convention for linking assigns
the second contribution with the opposite sign, because dragging the first
line around the second is inverse to dragging the second around the first.
Multiplying the two contributions gives the displayed phase.  The line
operators are mutually local precisely when the phase is one for all allowed
linking moves, which is \eqref{eq:thooft-mutual-locality-condition}.
\end{proof}

\paragraph{Screening equivalence of center-sensitive charges.}
Adjoint dynamical particles shift the electric weight of a line by an element
of the root lattice \(\Lambda_{\rm rt}\).  Smooth monopole cores shift the
magnetic cocharacter by an element of the coroot lattice
\(\Lambda_{\rm crt}\).  Therefore the finite charge of a genuine
Wilson--'t Hooft line is a class in
\[
  \bigl(\Lambda_{\rm wt}/\Lambda_{\rm rt}\bigr)
  \oplus
  \bigl(\Lambda_{\rm cwt}/\Lambda_{\rm crt}\bigr),
\]
subject to the mutual-locality condition imposed by the chosen theory.

An adjoint field transforms with weights in the root lattice together with
zero weights.  Attaching a massive adjoint particle worldline to a Wilson
line changes the line's electric weight by one of these roots.  The endpoint
of the adjoint worldline is a local charged insertion in the screened sector,
so the two Wilson lines differ by a locally creatable charge and have the
same finite one-form charge.

For the magnetic statement, a smooth finite-energy monopole embedded in the
\(\mathfrak{su}(2)\) subalgebra of a root \(\alpha\) has magnetic charge the
coroot \(\alpha^\vee\).  Bringing such a monopole worldline to the
't Hooft line changes the singular cocharacter by \(\alpha^\vee\) while
leaving the asymptotic bundle class modulo smooth monopole creation
unchanged.  The coroots generate \(\Lambda_{\rm crt}\), so magnetic charges
are screened modulo this lattice.

\subsection{Theta Angle and the Witten Effect}

In an abelian sector with curvature normalized by
\(\int_{\Sigma}F/(2\pi)\in\mathbb Z\), the theta term changes the electric
charge measured at infinity in a magnetic background.  With the convention
\[
  S_\theta=\frac{\ii\theta}{8\pi^2}\int_X F\wedge F
\]
in Euclidean signature, varying the action with respect to the boundary value
of \(A\) shows that the electric displacement receives the magnetic
contribution \((\theta/2\pi)m\).  Thus a line whose integral electric label
is \(\lambda\) and magnetic label is \(m\) has theta-dependent electric
charge
\[
  q_\theta=\lambda+\frac{\theta}{2\pi}m
\]
after pairing with any probe character.  Under \(\theta\mapsto\theta+2\pi\),
the integral label changes by the lattice automorphism
\begin{equation}
  T:(\lambda,m)\longmapsto(\lambda+m,m).
  \label{eq:witten-effect-T-action}
\end{equation}
For a nonabelian gauge theory the theta term contains a chosen invariant
integral quadratic form.  Its polarization gives a symmetric homomorphism
\[
  I:X_*(T)\longrightarrow X^*(T),
  \qquad
  I(m)(m')=I(m,m') .
\]
The Witten-effect automorphism is then
\[
  T_I:(\lambda,m)\longmapsto(\lambda+I(m),m).
\]
The abelian formula \eqref{eq:witten-effect-T-action} is the special case in
which \(I\) is the identity under the chosen electric-magnetic
identification.  The automorphism \(T_I\) preserves the Dirac pairing:
\[
  (\lambda+I(m))(m')-(\lambda'+I(m'))(m)
  =
  \lambda(m')-\lambda'(m),
\]
because \(I(m,m')=I(m',m)\).  Hence theta periodicity maps mutually local
line spectra to mutually local line spectra.

For \(G=U(1)\), the cocharacter lattice is \(\mathbb Z\).  The minimal
't Hooft line \(H(\gamma)\) has magnetic charge \(1\), so the boundary
condition is
\[
  \int_{S^2_p}\frac{F}{2\pi}=1
\]
for every small linking sphere.  If the line also carries electric charge
\(q\), one multiplies the magnetic boundary condition by a Wilson factor for
the unbroken gauge group along \(\gamma\); this gives a dyonic line.  A
dynamical monopole operator is an endpoint for such a magnetic line.  Hence a
theory containing dynamical monopoles does not have the magnetic one-form
symmetry under which closed 't Hooft lines are charged.

\section{Surface Operators}

A surface operator supported on a two-dimensional submanifold \(\Sigma\) may
be defined by singular behavior of the gauge field, by coupling a two-
dimensional QFT on \(\Sigma\) to the bulk, or by inserting a higher-form
charged object.  In four-dimensional gauge theory an abelianized surface
singularity has local model
\[
  A\sim \alpha\, d\varphi
\]
near the normal plane to \(\Sigma\), where \(\alpha\) is valued in a Cartan
subalgebra modulo the cocharacter lattice and Weyl group.  Additional
parameters can couple to the magnetic flux through \(\Sigma\).  A complete
definition must state the singularity, defect counterterms, gauge
equivalence, and how local operators approaching the surface are
renormalized.

Here is the corresponding local definition.  Let \(N_\Sigma\) be the
oriented real rank-two normal bundle of \(\Sigma\), and on each punctured
normal disk write polar coordinates \((r,\varphi)\).  A Cartan-type surface
singularity is a reduction of the gauge field near \(\Sigma\) to a maximal
torus \(T\) with
\begin{equation}
  A=\alpha\,\dd\varphi+A_{\rm reg},
  \qquad
  \alpha\in\mathfrak t,
  \label{eq:surface-alpha-singularity}
\end{equation}
where \(A_{\rm reg}\) has a finite limit as \(r\to0\) in the chosen regulator.
The holonomy around a small positively oriented normal circle is
\[
  \exp(2\pi\alpha)\in T
\]
after the convention that \(\mathfrak t\) is the compact real Lie algebra.
If \(u\in X_*(T)\), the gauge transformation
\[
  g_u(\varphi)=u(e^{\ii\varphi})
\]
shifts \(\alpha\) by \(u_*\).  Thus the singular holonomy parameter is
\[
  \alpha\in\mathfrak t/X_*(T)
\]
modulo the Weyl group.  In distributional notation the singular part of the
curvature is
\[
  F_{\rm sing}=2\pi \alpha\,\delta_\Sigma,
\]
where \(\delta_\Sigma\) is the two-form current Poincare dual to \(\Sigma\);
the formula is an instruction for the regulated boundary condition and for
the boundary values of correlation functions with the surface insertion.

An abelianized surface operator may also contain a two-dimensional theta
parameter
\[
  \eta\in X^*(T)\otimes_{\mathbb Z}\mathbb R/X^*(T),
\]
paired with magnetic flux through \(\Sigma\).  If \(F_T\) denotes the
regularized \(T\)-curvature on \(\Sigma\), the phase is
\[
  \exp\!\left(2\pi\ii\,\eta\!\left(\int_\Sigma\frac{F_T}{2\pi}\right)\right),
\]
with the flux interpreted in the cocharacter lattice.  The pair
\((\alpha,\eta)\), modulo cocharacter shifts and Weyl transformations, is the
center-sensitive datum of the abelianized surface singularity.  Coupling a
genuine two-dimensional QFT on \(\Sigma\) refines this datum by adding a
defect Hilbert space, a defect stress tensor, and defect counterterms.

\paragraph{Gauge equivalence of Cartan surface singularities.}
Two Cartan surface singularities with parameters \(\alpha,\alpha'\in
\mathfrak t\) have the same normal-circle holonomy datum exactly when
\[
  \alpha'=w(\alpha)+u_*
  \qquad
  \text{for some }w\in W,\; u\in X_*(T).
\]

A boundary-preserving gauge transformation may conjugate the chosen Cartan
reduction by an element of the normalizer \(N_G(T)\); modulo \(T\), this is
the Weyl action \(W=N_G(T)/T\).  A gauge transformation on the punctured
normal disk whose restriction to the normal circle is a map \(S^1\to T\) is
a cocharacter \(u\), and its logarithmic derivative is
\(u_*\,\dd\varphi\).  It shifts the coefficient of \(\dd\varphi\) by
\(u_*\).  These two operations generate the displayed equivalence.

Conversely, equal holonomies in the compact torus mean
\(\exp(2\pi\alpha')=\exp(2\pi w(\alpha))\) for some Weyl representative after
matching Cartan reductions.  The kernel of the exponential map
\(\mathfrak t\to T\) is the cocharacter lattice, so
\(\alpha'-w(\alpha)=u_*\) for some \(u\in X_*(T)\).

\section{Domain Walls and Interfaces}

A domain wall between theories \(T_-\) and \(T_+\) is a codimension-one defect
whose correlation functional takes bulk fields of \(T_-\) on one side and
bulk fields of \(T_+\) on the other.  A topological wall is one whose
correlators are invariant under deformations of the wall that avoid other
insertions.  Such a wall acts on local and extended operators by bringing an
operator to the wall and moving it through, provided the required limits
exist.

\begin{definition}[Fusion of topological defects]
\label{def:fusion-topological-defects}
Let \(D_1\) and \(D_2\) be parallel topological defects of the same
codimension.  Their fusion \(D_1\otimes D_2\), when it exists, is the defect
defined by the limit in which the separation between \(D_1\) and \(D_2\) goes
to zero, after subtracting or renormalizing local defect counterterms.
\end{definition}

Fusion is an operation only after the limiting procedure and equivalence
relation have been supplied.  In favorable topological sectors the resulting
objects form a category or higher category; in a general QFT the short-distance
limit can produce new defect degrees of freedom or fail to exist.

\begin{openproblem}[Extended-operator reconstruction]
\label{op:extended-operator-reconstruction}
Given the local operator algebra and symmetry data of a QFT, determine what
additional input is required to reconstruct its extended operators.  The
answer must include global forms, higher-form symmetries, singular boundary
conditions, defect renormalization, and the topology in which defect fusion
limits are taken.
\end{openproblem}
