\chapter{Boundaries And Defects}
\label{ch:global-boundaries-and-defects}

\ConventionsMixed

Boundaries and defects are QFT data supported on submanifolds.  They modify
operator algebras, symmetry realization, anomaly inflow, and phase structure.
This chapter formulates them as local data rather than as afterthoughts added
to a bulk theory.  The examples below are elementary enough that their fusion
and anomaly data can be checked directly.

\section{Boundary QFT Datum}

\begin{definition}[Boundary QFT datum]
\label{def:boundary-qft-datum}
Let \(M\) be a spacetime manifold with boundary \(\partial M\), and let
\(\mathcal A_{\rm bulk}\) be the local observable net on the interior
\(M^\circ\).  A boundary condition is an extension of this interior theory to
a boundary-preserving local net.  Choose a class
\(\operatorname{Reg}_\partial(M)\) of admissible regions in \(M\), including
interior regions, boundary regions, and collar regions meeting both strata.
The datum consists of:
\[
\begin{array}{ll}
\text{mixed-region observables} &
  \mathcal A_\partial(O),\quad O\in\operatorname{Reg}_\partial(M),\\
\text{interior identification} &
  \mathcal A_\partial(O)\cong\mathcal A_{\rm bulk}(O)
  \quad\text{if }O\Subset M^\circ,\\
\text{boundary observables} &
  \mathcal B(V),\quad V\subset\partial M,\\
\text{structure maps} &
  \iota_{O_1O_2}:\mathcal A_\partial(O_1)\to
  \mathcal A_\partial(O_2)
  \quad\text{for }O_1\subset O_2,\\
\text{boundary embeddings} &
  \jmath_{V,O}:\mathcal B(V)\to\mathcal A_\partial(O)
  \quad\text{when }V\subset O\cap\partial M .
\end{array}
\]
The maps are functorial under inclusions.  If the theory is Lorentzian, the
net is required to satisfy boundary-preserving covariance, causality, and the
time-slice axiom for globally hyperbolic regions with timelike boundary.  In
Euclidean or topological formulations the same slot is replaced by the
corresponding factorization products for disjoint subregions and collar
gluings.  Observables with spacelike separated supports commute or
graded-commute after their supports are interpreted in the manifold with
boundary.

A state, reflection-positive functional, or Euclidean amplitude is additional
data on this local algebraic system.  The boundary condition specifies the
allowed local observables and their structure maps before a state is chosen.
\end{definition}

Definition~\ref{def:boundary-qft-datum} deliberately contains no bare map
\(\mathcal A_{\rm bulk}(U)\to\mathcal B(V)\) attached to a symbol \(U\to V\).
An interior region \(U\) is included into a collar region, and a boundary
region \(V\) is included into the same collar region; these are ordinary
local-net structure maps.  Bringing a bulk local field to the boundary is a
short-distance operation.  When it exists, it is encoded by a renormalized
approach map or by a boundary operator expansion with a declared remainder
topology, as in Hypothesis~\ref{hyp:boundary-operator-expansion}.  This
separation is what lets boundary Ward identities, line endpoints, and anomaly
inflow be stated as local compatibility conditions rather than as an
untyped algebra homomorphism.

\section{Boundary Operator Expansion}

\begin{hypothesis}[Boundary operator expansion with remainder]
\label{hyp:boundary-operator-expansion}
For a scalar bulk local operator \(\mathcal O(x_\perp,y)\), with
\(x_\perp>0\) the distance to the boundary and \(y\in\partial M\), a
boundary-operator expansion datum consists of boundary local operators
\(\widehat{\mathcal O}_\alpha(y)\), coefficient distributions
\(c_{\mathcal O}^{\ \alpha}(x_\perp)\), an ordering of the retained boundary
operators, and a topology on smeared correlation functions with all other
insertions separated from \(y\).  The expansion has the form
\[
  \mathcal O(x_\perp,y)
  =
  \sum_\alpha
  c_{\mathcal O}^{\ \alpha}(x_\perp)\,\widehat{\mathcal O}_\alpha(y),
\]
in the following asymptotic sense.  For each cutoff \(N\), let
\[
  R_N(x_\perp,y)
  =
  \mathcal O(x_\perp,y)
  -
  \sum_{\alpha\leq N}
  c_{\mathcal O}^{\ \alpha}(x_\perp)\,\widehat{\mathcal O}_\alpha(y).
\]
The datum includes seminorms \(p_j\) and functions
\(\varepsilon_{j,N}(x_\perp)\to0\) such that
\[
  p_j\!\left(
    \langle R_N(x_\perp,y)\,X\rangle
  \right)
  \leq
  \varepsilon_{j,N}(x_\perp)
\]
for every allowed separated spectator insertion \(X\) in the stated class.
Stronger norm convergence or summability is an additional theorem or
BCFT-specific hypothesis.  Boundary critical exponents and boundary anomalies
are properties of this asymptotic expansion, the boundary stress tensor, and
the mixed-region local net.
\end{hypothesis}

The stronger two-dimensional BCFT package, where positive-energy boundary
Hilbert spaces, bulk-to-boundary expansion coefficients, and bordered-surface
sewing maps are part of the controlled datum, is recorded in
Chapter~\ref{ch:boundary-conformal-field-theory}.  The present chapter uses
the local-net and asymptotic-remainder slots above as the general boundary-QFT
interface needed for endpoints, fusion, and anomaly inflow.

\begin{example}[Free scalar half-space boundary expansion]
\label{ex:free-scalar-half-space-boe}
Let \(\phi\) be a free massless scalar field on the Euclidean half-space
\(\mathbb R_{\ge0}\times\mathbb R^{D-1}\), with coordinates
\((x_\perp,y)\).  For Neumann boundary condition
\(\partial_\perp\phi|_{x_\perp=0}=0\), the method of images gives
\[
  \langle\phi(x_\perp,y)\phi(x'_\perp,y')\rangle_N
  =
  G_D(x_\perp-x'_\perp,y-y')
  +
  G_D(x_\perp+x'_\perp,y-y').
\]
For Dirichlet boundary condition \(\phi|_{x_\perp=0}=0\), the sign of the
image term is reversed.  Consequently the leading boundary operator in the
Neumann case is \(\widehat\phi(y)=\phi(0,y)\), whereas in the Dirichlet case
the leading nonzero boundary operator is
\(\widehat{\partial_\perp\phi}(y)=\partial_\perp\phi(0,y)\).  This example
instantiates Definition~\ref{def:boundary-qft-datum}: interior smeared fields
and boundary smeared fields both map into collar algebras, while the displayed
two-point functions are Gaussian states on that local system.

For smooth test functions supported away from other insertions, the approach
to the boundary is an asymptotic statement.  In the Neumann case,
\[
  \phi(x_\perp,y)
  =
  \widehat\phi(y)+O(x_\perp^2),
\]
because the normal derivative vanishes in the boundary condition.  In the
Dirichlet case,
\[
  \phi(x_\perp,y)
  =
  x_\perp\,\widehat{\partial_\perp\phi}(y)+O(x_\perp^2),
\]
because the boundary value itself vanishes.  Evaluating the bulk field at
\(x_\perp=0\) would therefore erase the leading Dirichlet boundary operator;
the BOE is a renormalized approach datum, not an ordinary restriction map from
bulk algebras to boundary algebras.
\end{example}

\section{Interfaces And Domain Walls}

\begin{definition}[Interface and domain wall]
\label{def:interface-domain-wall}
An interface between two QFTs \(\mathcal T_-\) and \(\mathcal T_+\) is a
codimension-one defect whose local neighborhoods have \(\mathcal T_-\) on one
side and \(\mathcal T_+\) on the other.  A domain wall is an interface between
two phases or vacua of the same microscopic theory.  Topological interfaces
can be moved without changing separated correlation functions, up to contact
terms specified on the wall.  Non-topological interfaces carry ordinary
localized excitations and have their own stress-tensor Ward identities.
\end{definition}

\begin{example}[Chern--Simons boundary and chiral WZW currents]
\label{ex:cs-boundary-wzw-global-chapter}
For Chern--Simons theory on a three-manifold with boundary, the variation of
the bulk action contains the boundary one-form
\[
  \frac{k}{4\pi}\int_{\partial M}\operatorname{tr}(A\wedge\delta A).
\]
After choosing the holomorphic polarization on the boundary surface
\(\Sigma\), the boundary field \(g:\Sigma\to G\) has the chiral
Wess--Zumino--Witten action displayed in
Chapter~\ref{ch:tqft-chern-simons-theory}.  The boundary current
\[
  J(z)=-\frac{k}{2\pi}\partial_zg\,g^{-1}
\]
obeys the affine current algebra at level \(k\).  Thus the boundary condition
does not simply remove a boundary term; it creates a boundary operator
algebra whose representation theory is the state-space input for Wilson-line
endpoints and boundary conformal blocks.
\end{example}

\section{Fusion}

\begin{definition}[Renormalized defect fusion]
\label{def:renormalized-defect-fusion}
Let \(D_1\) and \(D_2\) be parallel defects separated by distance \(\ell\).
Fusion asks for the limit
\[
  D_2\circ_\ell D_1\quad \text{as}\quad \ell\to0 .
\]
The limit is a defect \(D_2\star D_1\) only if correlation functions with
separated external insertions converge after renormalizing defect-local
operators.  In a topological sector, fusion gives a monoidal category of
defects.  In a general QFT, fusion is a renormalization problem for
codimension-one local data.
\end{definition}

\begin{example}[Ising topological defect fusion]
\label{ex:ising-defect-fusion-global-boundaries}
At the critical two-dimensional Ising point, the basic topological defects
are
\[
  \mathbf 1,\qquad \eta,\qquad \mathcal N,
\]
where \(\eta\) is the invertible spin-flip defect and \(\mathcal N\) is the
Kramers--Wannier duality defect.  Their fusion rules are
\[
  \eta\otimes\eta\simeq\mathbf 1,\qquad
  \eta\otimes\mathcal N\simeq\mathcal N\otimes\eta\simeq\mathcal N,
  \qquad
  \mathcal N\otimes\mathcal N\simeq\mathbf 1\oplus\eta .
\]
The last equation is the first test that the word ``duality'' is being used
as a defect statement rather than as a slogan: \(\mathcal N\) has no inverse,
because its square is a direct sum.  Chapter~\ref{ch:global-categorical-symmetry-defect-fusion}
derives the corresponding action on primary sectors from the Ising modular
\(S\)-matrix.
\end{example}

\begin{remark}[Fusion detects noninvertibility]
\label{rem:ising-defect-noninvertible}
In the fusion algebra of Example~\ref{ex:ising-defect-fusion-global-boundaries},
the defect \(\mathcal N\) is not invertible and has Frobenius--Perron
dimension \(\sqrt2\).

If \(\mathcal N\) had an inverse \(X\), then
\(\mathcal N\otimes X\simeq\mathbf 1\) would be simple.  But fusing
\(\mathcal N\otimes\mathcal N\simeq\mathbf 1\oplus\eta\) with \(X\) gives
\[
  \mathcal N\simeq X\oplus(\eta\otimes X),
\]
which expresses the simple defect \(\mathcal N\) as a direct sum of two
nonzero summands, a contradiction in the semisimple defect category.  The
positive dimension homomorphism satisfies
\[
  d_{\mathcal N}^2=d_{\mathbf 1}+d_\eta=2,
\]
because \(d_{\mathbf 1}=d_\eta=1\).  Hence
\(d_{\mathcal N}=\sqrt2\).
\end{remark}

\begin{example}[Electric-magnetic duality surface in three-dimensional BF]
\label{ex:three-d-bf-duality-surface}
The \(\mathbb Z_N\) topological gauge theory can be written in continuum
notation as the \(U(1)\times U(1)\) BF action
\[
  S_{\rm BF}
  =
  \frac{N}{2\pi}\int B\wedge dA,
\]
with compact one-form gauge fields \(A\) and \(B\).  Its simple line labels
are pairs \((e,m)\in\mathbb Z_N\oplus\mathbb Z_N\), represented by
\[
  W_{e,m}(\gamma)
  =
  \exp\!\left(\ii e\int_\gamma A+\ii m\int_\gamma B\right).
\]
The codimension-one surface defect \(S\) that exchanges \(A\) and \(B\) acts
by
\[
  S:\ (e,m)\longmapsto(m,e).
\]
It is topological because it preserves the BF action and the braiding pairing
\[
  \langle(e,m),(e',m')\rangle
  =
  \exp\!\left(\frac{2\pi\ii}{N}(em'+me')\right).
\]
Fusion satisfies \(S\otimes S\simeq\mathbf 1\).  This is an invertible
surface defect; unlike the Ising \(\mathcal N\) defect, it is an ordinary
duality wall because the induced map on the finite line-label group is an
automorphism.
\end{example}

\section{Anomalies And Inflow}

\begin{definition}[Anomaly line of a boundary or defect]
\label{def:defect-anomaly-line}
A boundary or defect can carry an anomaly line for background fields
restricted to its support.  If \(A_D\) denotes those restricted background
fields, then a defect partition function with anomaly is a section
\[
  Z_D(A_D)\in\Gamma(\mathcal L_D),
\]
where \(\mathcal L_D\) is a complex line over the background-field space.
\end{definition}

An invertible bulk theory with boundary \(D\) supplies a section of
\(\mathcal L_D^{-1}\).  The product is then a complex number.  If a boundary
condition preserves a global symmetry, its boundary anomaly must match the
restriction of the bulk inflow.  If it breaks the symmetry, the broken
background fields are not part of the boundary datum.

\begin{remark}[Inflow cancellation as line trivialization]
\label{rem:inflow-line-trivialization}
The combined bulk-defect amplitude is an ordinary complex-valued functional
of background fields if and only if the tensor product of the defect anomaly
line and the bulk inflow line is canonically trivialized.

If \(Z_D(A_D)\) is a section of \(\mathcal L_D\) and the bulk inflow
amplitude \(Z_{\rm bulk}(A)\) is a section of \(\mathcal L_{\rm bulk}\), then
their product is a section of
\(\mathcal L_D\otimes\mathcal L_{\rm bulk}\).  It is a number precisely after
choosing a natural isomorphism
\[
  \mathcal L_D\otimes\mathcal L_{\rm bulk}\cong\underline{\C}.
\]
Conversely, if the product is a complex-valued functional for all background
fields and is compatible with gluing of backgrounds, it defines exactly such
a trivialization of the tensor product line.
\end{remark}

\section{Gauge-Theory Boundaries}

For gauge theories, a boundary condition must specify which gauge
transformations are redundancies at the boundary and which become global
symmetries acting on boundary operators.  Electric boundary conditions fix
the tangential connection or electric flux according to the chosen
Hamiltonian polarization.  Magnetic boundary conditions fix the dual data.
Line operators may end on the boundary precisely when boundary degrees of
freedom carry the corresponding charge and cancel the endpoint gauge
variation.

\begin{example}[Endpoint gauge variation of a Wilson line]
\label{ex:wilson-line-endpoint-boundary}
Let \(\gamma\) be an oriented path from \(p\) to \(q\), and let
\[
  U_\gamma(A)=\operatorname{Pexp}\int_\gamma A
\]
be parallel transport in a representation \(R\).  Under a gauge
transformation \(g\),
\[
  U_\gamma(A^g)=R(g(p))\,U_\gamma(A)\,R(g(q))^{-1}.
\]
Thus an open Wilson line is not gauge invariant by itself.  It can end on a
boundary only if boundary operators at \(p\) and \(q\) transform in the dual
representations needed to contract these endpoint factors, or if the
boundary condition converts the relevant gauge transformations into fixed
global symmetry transformations and the line is interpreted as a charged
operator under that boundary symmetry.
\end{example}

\section{Defects And Phase Data}

Boundary and defect categories refine the phase invariant of
Chapter~\ref{ch:global-phases-of-gauge-theories}.  A gapped phase includes
the fusion category of topological line defects, higher-codimension defect
operators, boundary conditions, and their anomaly responses.  Two theories
with the same local massive spectrum can be distinguished by these extended
data.

\begin{openproblem}[Defect-complete phase classification]
\label{op:defect-complete-phase-classification}
Construct a phase invariant that includes boundary conditions, defect fusion,
endpoint maps for line and surface operators, anomaly inflow, and local
operator data, and prove stability of this invariant under finite gapped
deformations.
\end{openproblem}
