\chapter{Vertex Operator Algebras, Modular Sewing, and Logarithmic CFT}
\label{ch:voa-modular-sewing-logarithmic-cft}

\ConventionsEuclidean

This chapter treats chiral algebra as the algebraic coordinate system for
two-dimensional CFT, but it keeps that coordinate system distinct from the
analytic construction of a full QFT on surfaces.  The first layer is formal:
VOA locality, modes, Zhu's algebra, minimal-model characters, Coulomb-gas
screenings, and logarithmic pseudo-traces are algebraic or finite-level
statements once the relevant modules are specified.  The second layer is
analytic within a fixed channel: chiral blocks, Gram/Shapovalov projectors,
trace-class propagation factors, and direct-integral OPE decompositions
require convergence, topological vector spaces, and positive pairings.  The
third layer is global sewing: local plumbing estimates have to be continued
across moduli space and compared between pants decompositions before they
become a modular functor or a full CFT.  The fourth layer is operator
algebraic: conformal nets and strongly local VOAs replace formal fields by
von Neumann algebras and make the domain-to-locality passage a theorem
boundary rather than a slogan.

The chapter's quoted-theorem boundaries are therefore not all of the same
kind.  Zhu classification and rational character modularity import finite
cofiniteness and trace-recursion theorems beyond the local top-level
calculations.  Complete rationality of conformal nets imports subfactor/DHR
machinery after the localized-endomorphism mechanism has been displayed.
Strong locality of VOAs imports analytic estimates turning smeared fields on a
common smooth domain into local von Neumann algebras.  Rational sewing imports
analytic continuation and mapping-class compatibility beyond fixed plumbing
charts.  Nonrational and logarithmic CFT require still different replacements
for finite semisimple sums: direct integrals, measurable fusing kernels,
projective pseudo-traces, modified traces, and nonsemisimple tensor-category
data.  The local calculations below are included to expose these mechanisms;
they are not meant to hide the remaining analytic existence problem.

\section{Chiral Algebra Data}

Two-dimensional conformal field theory has an additional algebraic structure:
local holomorphic fields may be multiplied by operator products that depend
only on one complex coordinate.  A vertex operator algebra is the precise
algebraic encoding of this chiral operator product.

\begin{definition}[Vertex operator algebra]
\label{def:vertex-operator-algebra}
A vertex operator algebra consists of the following data.
\begin{enumerate}
\item A complex vector space
\[
  V=\bigoplus_{n\in\mathbb Z}V_n
\]
with finite-dimensional graded pieces and \(V_n=0\) for \(n\) sufficiently
negative.
\item A vacuum vector \(\mathbf 1\in V_0\).
\item A state-field map
\[
  Y(\,\cdot\,,z):V\longrightarrow \operatorname{End}(V)[[z,z^{-1}]],
  \qquad
  a\longmapsto Y(a,z)=\sum_{m\in\mathbb Z}a_m z^{-m-1}.
\]
\item A conformal vector \(\omega\in V_2\), whose modes
\[
  Y(\omega,z)=\sum_{n\in\mathbb Z}L_n z^{-n-2}
\]
obey the Virasoro algebra
\[
  [L_m,L_n]=(m-n)L_{m+n}
  +\frac{c}{12}(m^3-m)\delta_{m+n,0}.
\]
\end{enumerate}
These data satisfy:
\[
  Y(\mathbf 1,z)=\operatorname{id}_V,\qquad
  Y(a,z)\mathbf 1\in a+zV[[z]],
\]
\[
  [L_{-1},Y(a,z)]=\partial_zY(a,z),
  \qquad
  L_0a=na\quad\text{for }a\in V_n,
\]
and locality:
for every \(a,b\in V\), there is \(N\ge0\) such that
\[
  (z-w)^N [Y(a,z),Y(b,w)]=0
\]
as a formal distribution on \(V\).
\end{definition}

The modes \(a_m\) are operators on \(V\).  If \(a\) has conformal weight
\(\Delta_a\), then \(a_m\) has degree \(\Delta_a-m-1\).  The vacuum axiom says
that \(a\) is recovered from the field \(Y(a,z)\) by acting on the vacuum and
then taking the value at \(z=0\):
\[
  a=\left. Y(a,z)\mathbf 1\right|_{z=0}.
\]
The locality axiom is the algebraic form of the chiral OPE.  It implies the
Jacobi identity for modes, and conversely the Jacobi identity packages
locality, associativity, and commutativity of OPE expansions.

\begin{proposition}[OPE from VOA locality]
\label{prop:voa-ope-from-locality}
For \(a,b\in V\), the product \(Y(a,z)Y(b,w)\) has an expansion in the region
\(|z|>|w|\) of the form
\[
  Y(a,z)Y(b,w)
  =
  \sum_{m\in\mathbb Z}Y(a_m b,w)(z-w)^{-m-1},
\]
where for each matrix element only finitely many positive powers of
\((z-w)^{-1}\) occur.
\end{proposition}

\begin{proof}
Fix \(a,b\in V\).  Lower truncation says that \(a_m b=0\) for all sufficiently
large \(m\); hence the singular part of the candidate expansion has only
finitely many negative powers of \(z-w\) in every matrix element.  Locality
gives an integer \(N\) such that
\[
  (z-w)^N Y(a,z)Y(b,w)=(z-w)^N Y(b,w)Y(a,z)
\]
as formal distributions.  The coefficient of \((z-w)^{-m-1}\) in the singular
part is determined by the formal residue
\[
  a_m b=\operatorname{Res}_{z-w}(z-w)^mY(a,z)b .
\]
For \(m<N\), these residues specify the complete principal part at \(z=w\).
Subtract the finite sum of principal parts from \(Y(a,z)Y(b,w)\).  After
multiplication by \((z-w)^N\), locality makes the difference regular at
\(z=w\), so the remaining term is the Taylor expansion in nonnegative powers
of \(z-w\) in the region \(|z|>|w|\).  Equivalently, all matrix elements
against graded-dual vectors agree as Laurent series in that region.  This is
the displayed OPE, and lower truncation gives the finiteness of its singular
part.
\end{proof}

\section{Modules, Characters, and Chiral Blocks}

A \(V\)-module \(M\) is a graded vector space with fields
\[
  Y_M(a,z)=\sum_{m\in\mathbb Z}a^M_m z^{-m-1}
\]
obeying the same vacuum, translation, and locality identities with one entry
acting on \(M\).  The conformal vector gives an operator \(L_0\) on \(M\).
For an ordinary positive-energy module,
\[
  M=\bigoplus_{\lambda\in\mathbb C}M_\lambda,\qquad
  \dim M_\lambda<\infty,
\]
and the set of \(\lambda\)'s is bounded below in every coset modulo
\(\mathbb Z\).  The character is
\begin{equation}
  \chi_M(\tau)
  =
  \operatorname{Tr}_M q^{L_0-c/24},
  \qquad
  q=\ee^{2\pi\ii\tau},
  \qquad
  \operatorname{Im}\tau>0 .
  \label{eq:voa-character}
\end{equation}
The shift by \(c/24\) is the cylinder vacuum energy fixed by the conformal map
from the plane to the circle.

Let \(C\) be a compact Riemann surface with marked points \(p_i\), local
coordinates \(z_i\) at the marked points, and \(V\)-modules \(M_i\).  A chiral
conformal block is a linear functional
\[
  \mathcal F:\bigotimes_i M_i\longrightarrow\mathbb C
\]
that obeys the Ward identities for meromorphic chiral fields on \(C\).  For
the stress tensor, the identity says that inserting a meromorphic vector field
regular away from the \(p_i\)'s acts by the sum of the corresponding Virasoro
modes at the marked points and gives zero.  For a general chiral state
\(a\in V\), the meromorphic one-form valued in the field \(a\) gives the same
condition with modes of \(Y_{M_i}(a,z_i)\).

On the sphere with three marked points, conformal blocks are intertwiners.
If \(M_i,M_j,M_k\) are simple modules, the fusion coefficient is
\[
  N_{ij}{}^k
  :=
  \dim \operatorname{CB}_{0,3}(M_i,M_j,M_k^\vee),
\]
where \(M_k^\vee\) is the contragredient module.  This definition is intrinsic:
it counts chiral three-point structures, not a particular coordinate
presentation of a three-point function.

\section{Zhu Algebra and Top Levels}

The character and sewing constructions use the whole graded module.  The
classification of ordinary modules begins with a smaller algebra obtained
from genus-one coordinate changes.  This algebra is Zhu's algebra.  Its role
in the monograph is structural: it converts the lowest-energy part of a
positive-energy module into a finite algebraic object, while leaving the
reconstruction of the full module to the VOA theorem boundary.

Let \(a\in V\) be homogeneous of conformal weight \(\Delta_a\), and write
\[
  Y(a,z)=\sum_{n\in\mathbb Z}a_n z^{-n-1}.
\]
Define bilinear operations on \(V\) by
\begin{align}
  a*b
  &:=
  \operatorname{Res}_{z=0}
  \frac{(1+z)^{\Delta_a}}{z}Y(a,z)b
  =
  \sum_{j\ge0}\binom{\Delta_a}{j}a_{j-1}b,
  \label{eq:zhu-star-product}\\
  a\circ b
  &:=
  \operatorname{Res}_{z=0}
  \frac{(1+z)^{\Delta_a}}{z^2}Y(a,z)b
  =
  \sum_{j\ge0}\binom{\Delta_a}{j}a_{j-2}b .
  \label{eq:zhu-circle-product}
\end{align}
The sums are finite on each pair \(a,b\) by lower truncation.  Let
\[
  O(V)=\operatorname{span}_{\mathbb C}\{a\circ b\mid a,b\in V
  \text{ homogeneous}\}.
\]

\begin{definition}[Zhu algebra]
\label{def:zhu-algebra}
The Zhu algebra of a VOA \(V\) is the quotient vector space
\begin{equation}
  A(V)=V/O(V)
  \label{eq:zhu-algebra-quotient}
\end{equation}
with product
\begin{equation}
  [a]\,[b]:=[a*b].
  \label{eq:zhu-product-on-classes}
\end{equation}
The formal identities below imply that this product is independent of the
chosen representatives and is associative; they are the same component-Jacobi
residue identities used in Proposition~\ref{prop:zhu-top-level-action}.  We
denote the class of \(a\in V\) by \([a]\in A(V)\).
\end{definition}

The exponent \((1+z)^{\Delta_a}\) is not decorative.  It is the coordinate
factor relating the punctured plane coordinate to the genus-one coordinate
used in torus one-point functions.  The quotient by \(O(V)\) removes the
descendant terms that vanish on lowest-energy states.

\begin{proposition}[Top-level action]
\label{prop:zhu-top-level-action}
Let \(M\) be an ordinary positive-energy \(V\)-module with lowest conformal
weight \(h_M\), so that
\[
  M=\bigoplus_{n\ge0}M_{h_M+n}.
\]
For a homogeneous \(a\in V\), set
\begin{equation}
  o_M(a):=a_{\Delta_a-1}\big|_{M_{h_M}} .
  \label{eq:zhu-zero-mode-action}
\end{equation}
Then \(o_M(a\circ b)=0\) on \(M_{h_M}\), and
\begin{equation}
  o_M(a*b)=o_M(a)o_M(b).
  \label{eq:zhu-top-product}
\end{equation}
Thus \(M_{h_M}\) is an \(A(V)\)-module.
\end{proposition}

\begin{proof}
The mode \(a_n\) sends an \(L_0\)-eigenvector of weight \(\lambda\) to weight
\begin{equation}
  \lambda+\Delta_a-n-1.
  \label{eq:voa-mode-weight-shift}
\end{equation}
Thus every lowest-weight vector \(v\in M_{h_M}\) is killed by \(u_n\) whenever
\(u\) is homogeneous and \(n>\Delta_u-1\).  Equivalently,
\begin{equation}
  M_{h_M}\subset
  \Omega(M):=\{v\in M\mid u_n v=0
  \text{ for all homogeneous }u\in V,\ n>\Delta_u-1\}.
  \label{eq:zhu-omega-subspace}
\end{equation}
The remaining step is a finite component calculation from the module Jacobi
identity.  In mode form, that identity says
\begin{align}
  \sum_{i\ge0}\binom{m}{i}(a_{n+i}b)_{m+k-i}
  &=
  \sum_{i\ge0}(-1)^i\binom{n}{i}
  \bigl(a_{m+n-i}b_{k+i}
  -(-1)^n b_{n+k-i}a_{m+i}\bigr).
  \label{eq:borcherds-component-identity}
\end{align}
All sums are finite on a fixed vector by lower truncation.  Substitute
\(n=j-2\) and sum with the coefficients
\(\binom{\Delta_a}{j}\) from \eqref{eq:zhu-circle-product}.  The resulting
finite sum is reorganized by the binomial identity
\((1+z)^{\Delta_a}=(1+z)(1+z)^{\Delta_a-1}\); every summand in
\(o_M(a\circ b)v\) contains a factor \(u_rv\) with \(r>\Delta_u-1\), hence
vanishes for \(v\in M_{h_M}\).  Substituting \(n=j-1\) and summing with the
coefficients in \eqref{eq:zhu-star-product} leaves exactly the term
\(a_{\Delta_a-1}b_{\Delta_b-1}v\); the remaining terms again contain a mode
killing \(\Omega(M)\).  Therefore
\begin{equation}
  o_M(a\circ b)v=0,\qquad
  o_M(a*b)v=o_M(a)o_M(b)v
  \label{eq:zhu-top-level-identities}
\end{equation}
for every \(v\in M_{h_M}\).  This proves that \(O(V)\) is in the kernel of
the top-level action and that the induced action is multiplicative.
\end{proof}

\begin{quotedtheorem}[Zhu classification theorem]
\label{qthm:zhu-classification}
Let \(V\) be a grading-restricted VOA of CFT type:
\[
  V_n=0\quad(n<0),\qquad V_0=\mathbb C\mathbf 1,\qquad
  \dim V_n<\infty .
\]
For every simple ordinary positive-energy \(V\)-module \(M\), the lowest
space \(M_{h_M}\) is a simple \(A(V)\)-module.  Conversely, every simple
\(A(V)\)-module \(U\) has a universal \(N\)-gradable weak \(V\)-module
generated from \(U\) as its top level; when the irreducible quotient is
ordinary, this construction is inverse to \(M\mapsto M_{h_M}\).  In
particular, if every simple \(A(V)\)-module induces an ordinary simple module,
then simple ordinary positive-energy \(V\)-modules are classified by simple
\(A(V)\)-modules.

If \(V\) is rational in the sense that the ordinary module category is
semisimple with finitely many simple objects and finite-dimensional graded
pieces, then \(A(V)\) is finite-dimensional and semisimple.  If, in addition,
\[
  C_2(V):=\operatorname{span}_{\mathbb C}\{a_{-2}b\mid a,b\in V\}
  \quad\text{has finite codimension in }V,
\]
then the same finite label set appears in the vector of characters
\((\chi_i(\tau))_{i\in I}\) entering the modular-invariance theorem.
This is the Zhu module-classification and character-finiteness theorem in the
form used here \cite{Zhu1996}.
\end{quotedtheorem}

\paragraph{Induction mechanism behind Zhu's theorem.}
The quoted theorem hides a concrete algebraic construction, so we record the
mechanism used later when examples are checked.  Let \(\mathfrak g(V)\) be
the Lie algebra generated by symbols \(a[m]\), for homogeneous \(a\in V\) and
\(m\in\mathbb Z\), modulo the translation relation
\[
  (L_{-1}a)[m]=-m\,a[m-1],
\]
with bracket
\begin{equation}
  [a[m],b[n]]
  =
  \sum_{j\ge0}\binom{m}{j}(a_jb)[m+n-j].
  \label{eq:zhu-mode-lie-algebra-bracket}
\end{equation}
The degree of \(a[m]\) is \(\Delta_a-m-1\).  Thus \(\mathfrak g(V)\) has a
triangular decomposition into positive, zero, and negative degree modes.  On
the top level of an ordinary module, positive degree modes vanish and zero
degree modes act by \(o(a)=a_{\Delta_a-1}\).  Proposition
\ref{prop:zhu-top-level-action} identifies this zero-degree action with the
Zhu algebra action.

Conversely, let \(U\) be a simple \(A(V)\)-module.  Give \(U\) an action of
the nonnegative part of \(\mathfrak g(V)\) by letting positive degree modes
act by zero and letting degree-zero modes act through \(A(V)\).  The induced
graded vector space
\begin{equation}
  \mathcal M_{\rm ind}(U)
  =
  U(\mathfrak g(V))\otimes_{U(\mathfrak g(V)_{\ge0})}U
  \label{eq:zhu-induced-module}
\end{equation}
is generated by \(U\) under negative degree modes.  The VOA Jacobi identity
imposes additional relations beyond the Lie brackets
\eqref{eq:zhu-mode-lie-algebra-bracket}; quotienting by those relations gives
the universal \(N\)-gradable weak module generated from \(U\).  Its maximal
proper graded submodule with zero intersection with \(U\), when it exists,
gives the irreducible quotient.  Every simple ordinary module is generated
from its top level by the negative modes, so it is obtained from this
construction.  The analytic and finiteness content of Quoted
Theorem~\ref{qthm:zhu-classification} is exactly the assertion that, under
the stated hypotheses, the quotient produced in this way is ordinary and that
the process exhausts the ordinary simple modules.

\begin{example}[Ising Zhu algebra as a top-weight test]
\label{ex:ising-zhu-algebra-top-weights}
For the Ising VOA \(L(1/2,0)\), the Zhu algebra is generated by the class
\([\omega]\) of the conformal vector.  The three simple modules have top
weights
\[
  0,\qquad \frac1{16},\qquad \frac12 .
\]
Equivalently, the semisimple Zhu algebra is the quotient
\begin{equation}
  A(L(1/2,0))
  \cong
  \mathbb C[x]\Big/
  \left(x\left(x-\frac1{16}\right)\left(x-\frac12\right)\right),
  \label{eq:ising-zhu-polynomial}
\end{equation}
where \(x\) acts as \(L_0\) on the top level.  The idempotents of this
three-dimensional algebra are the Lagrange polynomials
\begin{align}
  e_0
  &=
  32\left(x-\frac1{16}\right)\left(x-\frac12\right),
  \\
  e_{1/16}
  &=
  -\frac{256}{7}x\left(x-\frac12\right),
  \\
  e_{1/2}
  &=
  \frac{32}{7}x\left(x-\frac1{16}\right).
  \label{eq:ising-zhu-idempotents}
\end{align}
They obey \(e_h(h')=\delta_{h,h'}\), \(e_h^2=e_h\) in the quotient
\eqref{eq:ising-zhu-polynomial}, and \(e_0+e_{1/16}+e_{1/2}=1\).  Thus they
project to the vacuum, spin, and energy top levels, respectively.
\end{example}

\section{Unitary Virasoro Minimal Models}
\label{sec:unitary-virasoro-minimal-models}

The preceding VOA definitions are deliberately general.  The smallest
nontrivial unitary chiral CFTs arise when the chiral algebra is just the
Virasoro algebra and the representation category is finite.  This finiteness
is not automatic: it is a consequence of null vectors in highest-weight
Virasoro modules together with positivity.

Let \(M(c,h)\) be the Verma module generated by a highest-weight vector
\(\ket h\) satisfying
\[
  L_0\ket h=h\ket h,\qquad L_n\ket h=0\quad(n>0),
\]
for the Virasoro algebra of central charge \(c\).  In radial quantization of
a unitary chiral theory,
\[
  L_n^\dagger=L_{-n}.
\]
Thus each finite descendant level carries a Hermitian Gram form.  The first
two levels already show the mechanism.  At level one,
\[
  \|L_{-1}\ket h\|^2
  =
  \langle h|L_1L_{-1}|h\rangle
  =
  2h,
\]
so a unitary module has \(h\ge0\), and \(h=0\) makes \(L_{-1}\ket0\) a null
translation descendant of the vacuum.  At level two, in the ordered basis
\((L_{-2}\ket h,L_{-1}^2\ket h)\), the Gram matrix is
\begin{equation}
  M^{(2)}(c,h)
  =
  \begin{pmatrix}
    4h+\frac c2 & 6h\\
    6h & 4h(2h+1)
  \end{pmatrix}.
  \label{eq:virasoro-level-two-gram}
\end{equation}
Indeed,
\[
  \langle h|L_2L_{-2}|h\rangle
  =
  \langle h|(4L_0+c/2)|h\rangle,
\]
\[
  \langle h|L_2L_{-1}^2|h\rangle
  =
  6h,
\]
and
\[
  \langle h|L_1^2L_{-1}^2|h\rangle
  =
  4h(2h+1),
\]
obtained by commuting positive modes to the right until they annihilate
\(\ket h\).  If \(\det M^{(2)}(c,h)=0\) and \(h\ne0,-1/2\), the null vector is
\begin{equation}
  \left(
    L_{-2}-\frac{3}{2(2h+1)}L_{-1}^2
  \right)\ket h.
  \label{eq:level-two-virasoro-null-vector}
\end{equation}
Null descendants are not optional in a unitary CFT: they have zero inner
product with every state and must be quotiented to obtain the physical
irreducible module.

\begin{quotedtheorem}[Unitary Virasoro highest-weight classification]
\label{qthm:unitary-virasoro-minimal-classification}
Let \(L(c,h)\) be an irreducible positive-energy highest-weight module for the
Virasoro algebra with positive-definite Hermitian form and
\(L_n^\dagger=L_{-n}\).  Then either
\[
  c\ge1,\qquad h\ge0,
\]
or there is an integer \(m\ge3\) such that
\begin{equation}
  c=c_m:=1-\frac{6}{m(m+1)}
  \label{eq:unitary-minimal-central-charge}
\end{equation}
and
\begin{equation}
  h=h_{r,s}^{(m)}
  :=
  \frac{\bigl((m+1)r-ms\bigr)^2-1}{4m(m+1)}
  \label{eq:unitary-minimal-kac-weight}
\end{equation}
for integers
\[
  1\le r\le m-1,\qquad 1\le s\le m.
\]
The identification
\begin{equation}
  h_{r,s}^{(m)}=h_{m-r,m+1-s}^{(m)}
  \label{eq:kac-table-identification}
\end{equation}
leaves \(m(m-1)/2\) distinct irreducible modules.  The rational unitary
Virasoro minimal model \(\mathcal M(m,m+1)\) has precisely these simple
chiral modules.  This is the Friedan--Qiu--Shenker necessity theorem together
with the Goddard--Kent--Olive coset sufficiency theorem, stated here only at
the representation-classification boundary \cite{FriedanQiuShenker1984,
GoddardKentOlive1986}.
\end{quotedtheorem}

The theorem is the deep unitarity input.  The monograph uses it as a
classification boundary, not as a substitute for the local representation
calculations above.  The mechanism begins with the Kac determinant.  Introduce
a real parameter \(b>0\) by
\[
  c=1-6(b^{-1}-b)^2,
  \qquad
  h_{r,s}(b)
  =
  \frac{(r b^{-1}-s b)^2-(b^{-1}-b)^2}{4}.
\]
At descendant level \(N\), the determinant of the Virasoro Verma Gram form is,
up to a positive normalization depending on the chosen ordered monomial basis,
\begin{equation}
  \det M^{(N)}(c,h)
  =
  C_N
  \prod_{\substack{r,s\ge1\\rs\le N}}
  \bigl(h-h_{r,s}(b)\bigr)^{p(N-rs)},
  \qquad
  C_N>0,
  \label{eq:kac-determinant-unitary-mechanism}
\end{equation}
where \(p(k)\) is the number of integer partitions of \(k\).  Formula
\eqref{eq:kac-determinant-unitary-mechanism} is algebraic: it records exactly
where a singular vector can occur in a Verma module.  Unitarity then imposes an
inequality on every principal minor of every level Gram matrix.  For \(c<1\),
these inequalities are so restrictive that the zero hyperplanes
\(h=h_{r,s}(b)\) must occur in an interlacing pattern compatible with
quotienting null submodules.  The full Friedan--Qiu--Shenker/Goddard--Kent--Olive
argument proves that this happens precisely when
\[
  b^2=\frac{m}{m+1},
  \qquad
  m=3,4,\ldots,
\]
and the surviving irreducible highest weights are the Kac-table weights in
\eqref{eq:unitary-minimal-kac-weight}.  Thus the quoted theorem supplies the
all-level part of the classification: positivity of the null quotients at
every descendant level and exhaustion of the \(0<c<1\) positive-energy
highest-weight modules.

A convenient nonredundant list of representatives is the triangular set
\[
  1\le s\le r\le m-1.
\]

\paragraph{Rocha--Caridi characters.}
The next step uses one further representation-theoretic input: the
Rocha--Caridi alternating character formula for the irreducible quotients in
the unitary Kac table.  Its content is the Euler-character statement of the
singular-vector resolution of the irreducible Virasoro module.  Once this
character formula is accepted, the modular \(S\)-matrix below is a finite
theta-function calculation.

Put \(K=m(m+1)\), \(q=e^{2\pi\ii\tau}\), and
\[
  \Theta_{\lambda,K}(\tau)
  =
  \sum_{n\in\mathbb Z}
  q^{(2Kn+\lambda)^2/(4K)},
  \qquad
  \lambda\in\mathbb Z/2K\mathbb Z .
\]
For \(1\le r\le m-1\), \(1\le s\le m\), define
\[
  \lambda^-_{r,s}=(m+1)r-ms,
  \qquad
  \lambda^+_{r,s}=(m+1)r+ms .
\]
The irreducible minimal-model character is
\begin{equation}
  \chi_{r,s}^{(m)}(\tau)
  =
  \frac{
  \Theta_{\lambda^-_{r,s},K}(\tau)
  -
  \Theta_{\lambda^+_{r,s},K}(\tau)}
  {\eta(\tau)} .
  \label{eq:unitary-minimal-rocha-caridi-character}
\end{equation}

\begin{proposition}[Unitary minimal-model modular data from characters]
\label{prop:unitary-minimal-model-modular-data}
Let \(\mathcal M(m,m+1)\) be the diagonal unitary Virasoro minimal model
with \(m\ge3\), and assume that its irreducible chiral characters are the
Rocha--Caridi characters
\eqref{eq:unitary-minimal-rocha-caridi-character}.  Its chiral character
vector is indexed by Kac-table orbits
\[
  [r,s]=[m-r,m+1-s],
  \qquad
  1\le r\le m-1,\quad 1\le s\le m,
\]
or equivalently by the triangular representatives \(1\le s\le r\le m-1\).
The modular transformation \(\tau\mapsto-1/\tau\) is represented by
\begin{equation}
  S_{(r,s),(r',s')}
  =
  2\sqrt{\frac{2}{m(m+1)}}\,
  (-1)^{1+sr'+rs'}
  \sin\!\left(\frac{\pi(m+1)rr'}{m}\right)
  \sin\!\left(\frac{\pi mss'}{m+1}\right),
  \label{eq:unitary-minimal-model-s-matrix}
\end{equation}
and \(\tau\mapsto\tau+1\) is represented by the diagonal matrix
\begin{equation}
  T_{(r,s),(r,s)}
  =
  \exp\!\left(2\pi\ii\left(h_{r,s}^{(m)}-\frac{c_m}{24}\right)\right).
  \label{eq:unitary-minimal-model-t-matrix}
\end{equation}
The formula for \(S\) is well-defined on Kac-table orbits, is unitary on the
orbit basis, and together with \(T\) gives the genus-one modular
representation on characters.
\end{proposition}

\begin{proof}
The representation-theoretic content imported into the proposition is exactly
the character formula \eqref{eq:unitary-minimal-rocha-caridi-character}, whose
source is the Virasoro minimal-model character theory following the
Friedan--Qiu--Shenker and Goddard--Kent--Olive classification
\cite{FriedanQiuShenker1984,GoddardKentOlive1986}.  The modular matrix itself
is then computed as follows.  Poisson summation gives
\begin{equation}
  \Theta_{\lambda,K}(-1/\tau)
  =
  \sqrt{\frac{-\ii\tau}{2K}}\,
  \sum_{\mu\in\mathbb Z/2K\mathbb Z}
  e^{-\pi\ii\lambda\mu/K}\Theta_{\mu,K}(\tau),
  \qquad
  \eta(-1/\tau)=\sqrt{-\ii\tau}\,\eta(\tau).
  \label{eq:unitary-minimal-poisson-theta}
\end{equation}
Therefore
\[
  \chi_{r,s}^{(m)}(-1/\tau)
  =
  \frac{1}{\sqrt{2K}}
  \sum_{\mu\in\mathbb Z/2K\mathbb Z}
  \left(
  e^{-\pi\ii\lambda^-_{r,s}\mu/K}
  -
  e^{-\pi\ii\lambda^+_{r,s}\mu/K}
  \right)
  \frac{\Theta_{\mu,K}(\tau)}{\eta(\tau)} .
\]
Because \(\Theta_{\mu,K}=\Theta_{-\mu,K}\), the coefficient of
\(\Theta_{\lambda^-_{r',s'},K}/\eta\) after grouping the pair
\(\pm\lambda^-_{r',s'}\) is
\begin{align}
  &\frac{2}{\sqrt{2K}}
  \left[
  \cos\!\left(
  \frac{\pi\lambda^-_{r,s}\lambda^-_{r',s'}}{K}
  \right)
  -
  \cos\!\left(
  \frac{\pi\lambda^+_{r,s}\lambda^-_{r',s'}}{K}
  \right)
  \right]                                      \notag\\
  &\qquad =
  2\sqrt{\frac{2}{m(m+1)}}\,
  (-1)^{1+sr'+rs'}
  \sin\!\left(\frac{\pi(m+1)rr'}{m}\right)
  \sin\!\left(\frac{\pi mss'}{m+1}\right).
  \label{eq:unitary-minimal-poisson-s-coefficient}
\end{align}
The same calculation gives the negative of
\eqref{eq:unitary-minimal-poisson-s-coefficient} for
\(\Theta_{\lambda^+_{r',s'},K}/\eta\), so the coefficient of the character
\(\chi_{r',s'}^{(m)}\) is exactly
\eqref{eq:unitary-minimal-model-s-matrix}.  The \(T\)-matrix follows from
\(\Theta_{\lambda,K}(\tau+1)=e^{\pi\ii\lambda^2/(2K)}\Theta_{\lambda,K}(\tau)\),
\(\eta(\tau+1)=e^{\pi\ii/12}\eta(\tau)\), and
\((\lambda^+_{r,s})^2-(\lambda^-_{r,s})^2=4Krs\), which makes the two theta
terms in \eqref{eq:unitary-minimal-rocha-caridi-character} acquire the same
phase \(e^{2\pi\ii(h_{r,s}^{(m)}-c_m/24)}\).

The label identification is a quotient, not a mnemonic.  The involution
\[
  J(r,s)=(m-r,m+1-s)
\]
has no fixed points: a fixed point would require \(2r=m\) and \(2s=m+1\),
which cannot hold simultaneously in integers.  Thus every orbit has two
members, and the factor \(2\sqrt{2/(m(m+1))}\) in
\eqref{eq:unitary-minimal-model-s-matrix} is the quotient normalization.  If
one starts from the product of the two sine matrices for \(SU(2)_{m-2}\) and
\(SU(2)_{m-1}\), restriction to the \(J\)-orbits multiplies the parent
normalization by \(\sqrt2\).  This is the finite normalization step that is
often hidden in physics derivations.
\end{proof}

It is useful to state the resulting fusion rule without appealing to any
picture of fields.  Let \(n^{(k)}_{ab}{}^c\) be the \(SU(2)_k\) fusion
coefficient for \(0\le a,b,c\le k\):
\[
  n^{(k)}_{ab}{}^c=
  \begin{cases}
    1,& |a-b|\le c\le \min(a+b,2k-a-b)\ \text{and}\ a+b+c\in2\mathbb Z,\\
    0,& \text{otherwise}.
  \end{cases}
\]
For triangular representatives \(a=(r,s)\), \(b=(r',s')\), and
\(c=(r'',s'')\), the unitary minimal-model Verlinde coefficient is
\begin{equation}
\begin{split}
  N_{ab}{}^c
  &=
  n^{(m-2)}_{r-1,r'-1}{}^{r''-1}\,
  n^{(m-1)}_{s-1,s'-1}{}^{s''-1}
\\
  &\quad+
  n^{(m-2)}_{r-1,r'-1}{}^{m-r''-1}\,
  n^{(m-1)}_{s-1,s'-1}{}^{m-s''}.
\end{split}
  \label{eq:unitary-minimal-model-fusion-rule}
\end{equation}
The first term fuses to the displayed representative of the target orbit; the
second fuses to its \(J\)-image.

\begin{proposition}[Minimal-model fusion from the finite quotient]
\label{prop:unitary-minimal-model-fusion-check}
Assume the modular data of
Proposition~\ref{prop:unitary-minimal-model-modular-data}.  The Verlinde formula,
derived later in this chapter from modular diagonalization and displayed in
\eqref{eq:verlinde-formula}, gives exactly
\eqref{eq:unitary-minimal-model-fusion-rule}.  In particular the coefficients
are nonnegative integers.
\end{proposition}

\begin{proof}
Let
\[
  S^{(k)}_{ab}=\sqrt{\frac{2}{k+2}}\,
  \sin\!\left(\frac{\pi(a+1)(b+1)}{k+2}\right)
\]
be the \(SU(2)_k\) sine matrix.  The finite sine orthogonality relation
\[
  \sum_{x=0}^k
  S^{(k)}_{ax}S^{(k)}_{bx}=\delta_{ab}
\]
is the ordinary orthogonality of sine functions on the finite interval.
Put
\[
  \theta_x=\frac{\pi(x+1)}{k+2},
  \qquad
  \chi_a(\theta)=\frac{\sin((a+1)\theta)}{\sin\theta}.
\]
Then \(S^{(k)}_{ax}=S^{(k)}_{0x}\chi_a(\theta_x)\), and the Verlinde
coefficient is the finite inner product
\begin{equation}
  n^{(k)}_{ab}{}^c
  =
  \sum_{x=0}^k
  \bigl(S^{(k)}_{0x}\bigr)^2\,
  \chi_a(\theta_x)\chi_b(\theta_x)\chi_c(\theta_x).
  \label{eq:su2-finite-character-inner-product}
\end{equation}
The ordinary \(SU(2)\) character identity gives
\begin{equation}
  \chi_a\chi_b
  =
  \sum_{\substack{|a-b|\le d\le a+b\\ a+b+d\ {\rm even}}}\chi_d .
  \label{eq:ordinary-su2-character-product}
\end{equation}
On the finite grid \(\theta_x\) there are the affine-wall relations
\begin{equation}
  \chi_{k+1}(\theta_x)=0,\qquad
  \chi_{k+1+r}(\theta_x)=-\chi_{k+1-r}(\theta_x)
  \quad (r\ge1),
  \label{eq:su2-affine-wall-character-relation}
\end{equation}
which follow from \((k+2)\theta_x=\pi(x+1)\).  Substituting
\eqref{eq:ordinary-su2-character-product} in
\eqref{eq:su2-finite-character-inner-product}, reducing all terms by
\eqref{eq:su2-affine-wall-character-relation}, and using finite
orthogonality gives
\begin{equation}
  n^{(k)}_{ab}{}^c
  =
  \begin{cases}
  1,& |a-b|\le c\le \min(a+b,2k-a-b),\quad a+b+c\in2\mathbb Z,\\
  0,& \text{otherwise}.
  \end{cases}
  \label{eq:su2k-fusion-from-finite-grid}
\end{equation}
Indeed, if \(a+b\le k\) no wall term occurs and this is the ordinary
\(SU(2)\) tensor-product interval.  If \(a+b>k\), the reflected terms across
the wall \(k+1\) cancel exactly the labels \(c>2k-a-b\), leaving the truncated
alcove interval displayed above.

For the minimal model, the matrix
\eqref{eq:unitary-minimal-model-s-matrix} is the signed product of the
\(SU(2)_{m-2}\) and \(SU(2)_{m-1}\) sine matrices, followed by the
fixed-point-free quotient by \(J\).  The sign is the simple-current selection
rule in the Verlinde sum.  Therefore the quotient coefficient is obtained from
the parent product by summing over the two representatives \(c\) and \(Jc\) of
the target orbit.  Written in triangular labels, this gives exactly
\eqref{eq:unitary-minimal-model-fusion-rule}.  Each summand is a product of
\(0\)-\(1\) \(SU(2)_k\) fusion coefficients, so the result is a nonnegative
integer.
\end{proof}

\subsection{Coulomb-Gas Blocks and Screening Charges}
\label{subsec:minimal-model-coulomb-gas}

There is a second, complementary description of Virasoro minimal-model chiral
blocks: a free chiral scalar with background charge, together with contour
integrals of screening fields.  This construction is useful only after its
domain of validity is stated carefully.  It constructs chiral blocks and
intertwining maps in a free-field resolution of Virasoro modules; it is not by
itself a definition of the full nonchiral CFT, nor does it replace the Hilbert
space positivity and sewing data discussed above.

Let \(\varphi(z)\) be a chiral scalar normalized by
\begin{equation}
  \varphi(z)\varphi(w)\sim-\log(z-w),
  \qquad
  \partial\varphi(z)\partial\varphi(w)\sim-\frac1{(z-w)^2}.
  \label{eq:coulomb-gas-free-scalar-ope}
\end{equation}
For a complex background-charge parameter \(\alpha_0\), define
\begin{equation}
  T_{\alpha_0}(z)
  =
  -\frac12:\partial\varphi\,\partial\varphi:(z)
  +\ii\alpha_0\partial^2\varphi(z).
  \label{eq:coulomb-gas-stress-tensor}
\end{equation}
For \(V_\alpha(z)=:e^{\ii\alpha\varphi(z)}:\), Wick contraction gives
\begin{equation}
  T_{\alpha_0}(z)V_\alpha(w)
  \sim
  \frac{h(\alpha)V_\alpha(w)}{(z-w)^2}
  +
  \frac{\partial V_\alpha(w)}{z-w},
  \qquad
  h(\alpha)=\frac12\alpha(\alpha-2\alpha_0),
  \label{eq:coulomb-gas-vertex-weight}
\end{equation}
and
\begin{equation}
  c=1-12\alpha_0^2.
  \label{eq:coulomb-gas-central-charge}
\end{equation}
The sign of the improvement term in
\eqref{eq:coulomb-gas-stress-tensor} is fixed by the convention
\eqref{eq:coulomb-gas-free-scalar-ope}; changing the sign of
\(\varphi\) changes \(\alpha\) and \(\alpha_0\) together and leaves
\eqref{eq:coulomb-gas-vertex-weight} invariant.

For the unitary model \(\mathcal M(m,m+1)\), set
\begin{equation}
  \alpha_+^2=\frac{2(m+1)}{m},
  \qquad
  \alpha_-^2=\frac{2m}{m+1},
  \qquad
  \alpha_+\alpha_-=-2,
  \qquad
  \alpha_0=\frac{\alpha_++\alpha_-}{2}.
  \label{eq:coulomb-gas-unitary-alpha}
\end{equation}
Thus \(m\alpha_+ +(m+1)\alpha_-=0\).  The charge associated with the Kac label
\((r,s)\) is
\begin{equation}
  \alpha_{r,s}
  =
  \frac{1-r}{2}\alpha_+
  +
  \frac{1-s}{2}\alpha_- .
  \label{eq:coulomb-gas-kac-charge}
\end{equation}
Substitution into \eqref{eq:coulomb-gas-central-charge} and
\eqref{eq:coulomb-gas-vertex-weight} gives
\begin{equation}
  c=1-\frac{6}{m(m+1)}=c_m,
  \qquad
  h(\alpha_{r,s})
  =
  \frac{\bigl((m+1)r-ms\bigr)^2-1}{4m(m+1)}
  =
  h_{r,s}^{(m)}.
  \label{eq:coulomb-gas-kac-weight}
\end{equation}
The Kac reflection is charge conjugation in the background-charge scalar:
\[
  \alpha_{m-r,m+1-s}=2\alpha_0-\alpha_{r,s},
  \qquad
  h(2\alpha_0-\alpha)=h(\alpha).
\]

\begin{definition}[Screening currents and Coulomb-gas block datum]
\label{def:coulomb-gas-screening-block-datum}
The screening currents are
\begin{equation}
  S_\pm(u)=V_{\alpha_\pm}(u).
  \label{eq:coulomb-gas-screening-currents}
\end{equation}
Because \(\alpha_+\alpha_-=-2\), \eqref{eq:coulomb-gas-vertex-weight} gives
\(h(\alpha_+)=h(\alpha_-)=1\).  Given external charges
\(\beta_i\), insertion points \(z_i\), and nonnegative integers
\(N_\pm\), remove from
\(\mathbb C^{N_++N_-}\) all diagonals \(u_a=u_{a'}\),
\(v_b=v_{b'}\), \(u_a=v_b\), and all collision hyperplanes with the
external points \(z_i\).  The multivalued factor in
\eqref{eq:coulomb-gas-block-integral} determines a rank-one local system
\(\mathcal L\) on this complement; a twisted cycle \(\Gamma\) means a
singular chain with coefficients in \(\mathcal L^\vee\) whose twisted
boundary vanishes.  Endpoint prescriptions in a compactification, when used,
are part of the choice of this relative twisted homology class.  A
genus-zero Coulomb-gas block datum consists of such a twisted cycle
\(\Gamma\) and the neutrality condition
\begin{equation}
  \sum_i\beta_i+N_+\alpha_+ +N_-\alpha_- =2\alpha_0.
  \label{eq:coulomb-gas-neutrality}
\end{equation}
The corresponding chiral integral is
\begin{equation}
\begin{split}
  \mathcal F_\Gamma(z_i)
  &=
  \int_\Gamma
  \prod_{a=1}^{N_+}\dd u_a
  \prod_{b=1}^{N_-}\dd v_b\,
  \prod_{i<j}(z_i-z_j)^{\beta_i\beta_j}
\\
  &\quad\times
  \prod_{a,i}(u_a-z_i)^{\alpha_+\beta_i}
  \prod_{b,i}(v_b-z_i)^{\alpha_-\beta_i}
  \prod_{a<a'}(u_a-u_{a'})^{\alpha_+^2}
  \prod_{b<b'}(v_b-v_{b'})^{\alpha_-^2}
  \prod_{a,b}(u_a-v_b)^{-2}.
\end{split}
  \label{eq:coulomb-gas-block-integral}
\end{equation}
\end{definition}

\begin{lemma}[Screenings produce Virasoro intertwiners]
\label{lem:coulomb-gas-screening-intertwiners}
For any closed contour or twisted cycle \(\Gamma\) on which boundary terms
vanish, insertion of the screening charges
\[
  Q_\pm(\Gamma)=\int_\Gamma S_\pm(u)\,\dd u
\]
commutes with the Virasoro action generated by \(T_{\alpha_0}\).  Consequently
\eqref{eq:coulomb-gas-block-integral}, when convergent or defined by analytic
continuation from a convergent chamber, satisfies the Virasoro Ward identities
for a chiral block with external weights \(h(\beta_i)\).
\end{lemma}

\begin{proof}
First suppose that \(\Gamma\) is represented by a compact twisted chain in a
convergence chamber, so that all contour deformations and integrations by
parts below are ordinary absolutely convergent operations after choosing local
branches.  Since \(S_\pm\) have weight \(1\),
\[
  T(z)S_\pm(u)
  \sim
  \frac{S_\pm(u)}{(z-u)^2}
  +
  \frac{\partial_u S_\pm(u)}{z-u}.
\]
The commutator of \(L_n=\oint z^{n+1}T(z)\dd z/(2\pi\ii)\) with the local
field \(S_\pm(u)\) is therefore
\[
  [L_n,S_\pm(u)]
  =
  \partial_u\!\left(u^{n+1}S_\pm(u)\right).
\]
Thus
\[
  [L_n,Q_\pm(\Gamma)]
  =
  \int_\Gamma
  d_u\!\left(u^{n+1}S_\pm(u)\right),
\]
where the differential is read as the flat covariant derivative on the
rank-one local system carried by the multivalued integrand.  Twisted Stokes
gives
\[
  \int_\Gamma \nabla_{\mathcal L}\eta
  =
  \int_{\partial_{\mathcal L}\Gamma}\eta .
\]
The right hand side vanishes because \(\partial_{\mathcal L}\Gamma=0\).  In
the single-valued special case this is the usual statement that the integral
of a total derivative around a closed contour is zero; in the multivalued
case the coefficients of the twisted chain are exactly what cancels the
monodromy terms on adjacent boundary faces.  For several screenings the same
argument is applied to each screening coordinate, and the commutator with
the product of screening integrations is the sum of these twisted exact
terms.  Hence the integrated screening charges commute with all Virasoro
modes in the sense of chiral blocks.

It remains to identify the Ward identity obeyed by the unintegrated free
field correlator.  Wick contraction with
\eqref{eq:coulomb-gas-stress-tensor} gives the primary terms
\eqref{eq:coulomb-gas-vertex-weight} at each finite exponential insertion.
The contour of \(\partial\varphi\) around all finite exponential insertions
measures the total finite charge
\(\sum_i\beta_i+N_+\alpha_+ +N_-\alpha_-\).  The improvement term in
\(T_{\alpha_0}\) is equivalently a background charge \(2\alpha_0\) at
infinity on the sphere.  Therefore the stress-tensor contour at infinity has
the vacuum behavior needed for the genus-zero Virasoro Ward identity exactly
when \eqref{eq:coulomb-gas-neutrality} holds.  The screening integrations
preserve that identity because their Virasoro commutators are the twisted
exact terms just shown to integrate to zero.

When the displayed integral is defined by meromorphic continuation, the Ward
identity is first proved in a nonempty convergence chamber with the same
twisted homology class.  Both sides are meromorphic functions of the charges
and insertion positions away from collision and resonance divisors, so the
identity extends to the continued block on the common domain of definition.
\end{proof}

\paragraph{One-screening Dotsenko--Fateev beta integral.}
Let \(A,B\in\mathbb C\) with initially
\(\operatorname{Re}A>-1\), \(\operatorname{Re}B>-1\), and choose the principal
branch on \((0,1)\).  Then the one-screening integral
\begin{equation}
  I(A,B)=\int_0^1 u^A(1-u)^B\,\dd u
  \label{eq:dotsenko-fateev-beta-integral}
\end{equation}
is
\begin{equation}
  I(A,B)
  =
  \frac{\Gamma(A+1)\Gamma(B+1)}
       {\Gamma(A+B+2)}.
  \label{eq:dotsenko-fateev-beta-result}
\end{equation}
Both sides define the meromorphic continuation of the same chiral block
coordinate in \(A,B\), with poles exactly where endpoint subtractions or
resonant descendants have to be specified.

In the convergence chamber this is Euler's beta integral: write
\[
  \Gamma(A+1)\Gamma(B+1)
  =
  \int_0^\infty\!\!\int_0^\infty
  e^{-x-y}x^A y^B\,\dd x\,\dd y.
\]
Use the change of variables \(t=x+y\), \(u=x/(x+y)\).  The Jacobian is
\(t\), so the double integral factors as
\[
  \left(\int_0^\infty e^{-t}t^{A+B+1}\dd t\right)
  \left(\int_0^1u^A(1-u)^B\dd u\right)
  =
  \Gamma(A+B+2)I(A,B).
\]
Division by \(\Gamma(A+B+2)\) gives
\eqref{eq:dotsenko-fateev-beta-result}.  Meromorphic continuation follows
because the gamma-function expression is meromorphic and agrees with the
integral on a connected open convergence chamber.

\paragraph{Two-screening Selberg chamber.}
The first place where the Dotsenko--Fateev integrals are not just repeated
Euler beta integrals is the two-screening factor.  In a real convergence
chamber, suppressing the common external-position prefactors, put
\begin{equation}
  I_2(A,B,C)
  =
  \int_0^1\!\!\int_0^1
  u^A(1-u)^Bv^A(1-v)^B|u-v|^{2C}\,\dd u\,\dd v .
  \label{eq:dotsenko-fateev-two-screening-selberg-integral}
\end{equation}
For \(\operatorname{Re}A,\operatorname{Re}B>-1\) and
\[
  \operatorname{Re}C
  >
  -\min\{1/2,\operatorname{Re}(A+1),\operatorname{Re}(B+1)\},
\]
Selberg's \(n=2\) formula gives
\begin{equation}
  I_2(A,B,C)
  =
  \frac{
    \Gamma(A+1)\Gamma(B+1)
    \Gamma(A+C+1)\Gamma(B+C+1)\Gamma(2C+1)
  }{
    \Gamma(A+B+C+2)
    \Gamma(A+B+2C+2)
    \Gamma(C+1)
  } .
  \label{eq:dotsenko-fateev-two-screening-selberg-result}
\end{equation}
For nonnegative integer \(A,B,C\) the same identity is already visible by
elementary finite algebra.  By symmetry,
\[
  I_2(A,B,C)
  =
  2\int_{0<v<u<1}
  u^A(1-u)^Bv^A(1-v)^B(u-v)^{2C}\,\dd v\,\dd u .
\]
Expand the three polynomials
\[
  (1-u)^B,\qquad (1-v)^B,\qquad (u-v)^{2C}.
\]
The term with powers \(i,j,k\) contributes
\begin{equation}
  2(-1)^{i+j+k}
  \binom{B}{i}\binom{B}{j}\binom{2C}{k}
  \frac{1}{A+j+k+1}
  \frac{1}{2A+i+j+2C+2}.
  \label{eq:dotsenko-fateev-two-screening-finite-sum}
\end{equation}
Summing \eqref{eq:dotsenko-fateev-two-screening-finite-sum} over
\(0\le i,j\le B\) and \(0\le k\le2C\) gives exactly
\eqref{eq:dotsenko-fateev-two-screening-selberg-result} with gamma functions
replaced by factorials.  Thus the new datum in the two-screening block is
not an additional copy of the one-screening endpoint powers; it is the
screening--screening exponent \(2C\), which is \(\alpha_\pm^2\) in the
Coulomb-gas block and is the source of the Selberg/Vandermonde factors in
multi-screening Dotsenko--Fateev formulas.

\paragraph{Screening-to-BPZ construction budget.}
The preceding formulas are coordinates on a block space; they are not yet a
proof that a chosen contour family is the correct minimal-model correlator.
On a finite jet test space in the cross-ratio, one should separate the
following finite vectors:
\[
  \mathcal C_{\rm DF},\quad
  \mathcal C_{\rm Ward},\quad
  \mathcal C_{\rm tw},\quad
  \mathcal C_{\rm BPZ},\quad
  \mathcal C_{\rm cont},\quad
  \mathcal C_{\rm full}.
\]
They denote, respectively, the raw Dotsenko--Fateev screening integrals, the
same coordinates after the Virasoro Ward/intertwiner check, the projection to
closed twisted cycles, the comparison with a BPZ solution basis, the analytically
continued block coordinates in a second channel, and the nonchiral sewing
pairing.  If \(\mathsf M_{\rm cont}\) is the finite connection matrix between
the chosen BPZ bases, define
\begin{align}
  E_{\rm scr}
  &=
  \mathcal C_{\rm Ward}-\mathcal C_{\rm DF},
  &
  E_{\rm tw}
  &=
  \mathcal C_{\rm tw}-\mathcal C_{\rm Ward},
  \label{eq:coulomb-gas-bpz-budget-a}\\
  E_{\rm BPZ}
  &=
  \mathcal C_{\rm BPZ}-\mathcal C_{\rm tw},
  &
  E_{\rm cont}
  &=
  \mathcal C_{\rm cont}-\mathsf M_{\rm cont}\mathcal C_{\rm BPZ},
  \label{eq:coulomb-gas-bpz-budget-b}\\
  E_{\rm pair}
  &=
  \mathcal C_{\rm full}-\mathcal C_{\rm cont}.
  \label{eq:coulomb-gas-bpz-budget-c}
\end{align}
Then the finite comparison identity is
\begin{equation}
  \mathcal C_{\rm full}-\mathsf M_{\rm cont}\mathcal C_{\rm DF}
  =
  \mathsf M_{\rm cont}
  (E_{\rm scr}+E_{\rm tw}+E_{\rm BPZ})
  +E_{\rm cont}+E_{\rm pair}.
  \label{eq:coulomb-gas-bpz-budget-telescope}
\end{equation}
Indeed, add and subtract the intermediate coordinates before applying the
connection matrix.  If the norm of \(\mathsf M_{\rm cont}\) on this finite test
space is at most \(\kappa\), and if the five residuals are bounded by
\(\varepsilon_{\rm scr},\varepsilon_{\rm tw},\varepsilon_{\rm BPZ},
\varepsilon_{\rm cont},\varepsilon_{\rm pair}\), then
\begin{equation}
  \|\mathcal C_{\rm full}-\mathsf M_{\rm cont}\mathcal C_{\rm DF}\|
  \leq
  \kappa(\varepsilon_{\rm scr}+\varepsilon_{\rm tw}
  +\varepsilon_{\rm BPZ})
  +\varepsilon_{\rm cont}+\varepsilon_{\rm pair}.
  \label{eq:coulomb-gas-bpz-budget-bound}
\end{equation}
The residuals have distinct meanings:
\begin{itemize}
\item \(E_{\rm scr}\) measures failure of the screening insertion to satisfy the
Ward/intertwiner relation in the declared regularization.
\item \(E_{\rm tw}\) measures failure of the contour to be a closed twisted
cycle, including endpoint and relative-boundary prescriptions.
\item \(E_{\rm BPZ}\) measures the finite BPZ differential-equation residual
after quotienting by null vectors.
\item \(E_{\rm cont}\) measures analytic continuation and connection-matrix
error between the two chosen channel bases.
\item \(E_{\rm pair}\) measures the nonchiral reflection-positive pairing and
sewing residual.
\end{itemize}
Thus a Dotsenko--Fateev formula becomes a theorem-level minimal-model block
construction only when these five residuals vanish compatibly as the finite jet
space is enlarged.

\begin{remark}[What the Dotsenko--Fateev formulas do and do not prove]
\label{rem:dotsenko-fateev-status}
Multiple screening integrals are Selberg-type generalizations of
\eqref{eq:dotsenko-fateev-beta-integral}.  Their contour choices span spaces
of solutions to the BPZ differential equations and are a powerful way to
compute chiral-block connection matrices and selected structure constants.
The load-bearing data are the charges, the neutrality equation, the twisted
cycle, and the analytic continuation prescription.  A single displayed
integral, without those data and without the nonchiral sewing pairing, is not
a construction of the full minimal model.  Conversely, once the Hilbert space,
null quotient, modular data, and sewing constraints have been fixed, the
Coulomb-gas integrals give concrete coordinates on the same chiral block
spaces and make many BPZ solutions explicitly computable.
\end{remark}

Consequently the diagonal A-series full CFT has genus-one partition function
\[
  Z_m(\tau,\bar\tau)
  =
  \sum_{1\le s\le r\le m-1}
  \left|\chi_{r,s}(\tau)\right|^2,
\]
which is modular invariant because the same unitary \(S\) and diagonal \(T\)
act on the holomorphic and antiholomorphic character vectors.

For \(m=3\), one obtains
\[
  c_3=\frac12,\qquad
  h_{1,1}=0,\qquad h_{2,1}=\frac12,\qquad h_{2,2}=\frac1{16}.
\]
The same spin field may also be represented as \(h_{1,2}=1/16\), which is the
label making its level-two null vector manifest.  For
\(c=1/2\) and \(h=1/16\), the matrix
\eqref{eq:virasoro-level-two-gram} has determinant zero and
\eqref{eq:level-two-virasoro-null-vector} becomes
\begin{equation}
  \left(L_{-2}-\frac43 L_{-1}^2\right)\ket\sigma=0.
  \label{eq:ising-sigma-level-two-null}
\end{equation}

\begin{proposition}[Ising spin four-point function from the BPZ equation]
\label{prop:ising-spin-bpz-four-point}
Normalize the Ising spin field by
\(\langle\sigma(z,\bar z)\sigma(0)\rangle=|z|^{-1/4}\).  Let
\[
  \langle\sigma(z_1,\bar z_1)\cdots\sigma(z_4,\bar z_4)\rangle
  =
  |z_{12}|^{-1/4}|z_{34}|^{-1/4}f(z,\bar z)
\]
in the usual four-point cross-ratio frame.  The null vector
\eqref{eq:ising-sigma-level-two-null} implies the holomorphic BPZ equation
\begin{equation}
  \left[
    \partial_z^2
    +\frac{2-5z}{4z(1-z)}\partial_z
    -\frac{3}{64(1-z)^2}
  \right]f(z,\bar z)=0,
  \label{eq:ising-sigma-bpz-equation}
\end{equation}
and the analogous antiholomorphic equation.  A basis of local holomorphic
solutions is
\[
  f_\pm(z)=(1-z)^{-1/8}\sqrt{1\pm\sqrt z}.
\]
If
\[
  \mathcal F_0(z)=\frac{f_+(z)+f_-(z)}2,\qquad
  \mathcal F_{\varepsilon}(z)=f_+(z)-f_-(z),
\]
then single-valuedness, normalization, and crossing fix
\begin{equation}
  f(z,\bar z)
  =
  |\mathcal F_0(z)|^2+\frac14|\mathcal F_{\varepsilon}(z)|^2,
  \qquad
  C_{\sigma\sigma\varepsilon}=\frac12
  \label{eq:ising-spin-four-point-structure-constant}
\end{equation}
up to the sign convention for \(\varepsilon\).
\end{proposition}

\begin{proof}
The BPZ equation follows from the null vector as follows.  Insert
\eqref{eq:ising-sigma-level-two-null} at one point in the four-point
function.  The \(L_{-1}^2\) term is the second derivative with respect to
that point.  The \(L_{-2}\) term is computed by writing \(L_{-2}\) as a
small contour integral of \(T(\zeta)\), expanding the contour to surround the
other three insertions, and using the primary Ward identity
\[
  T(\zeta)\sigma(z_i)
  \sim
  \frac{(1/16)\sigma(z_i)}{(\zeta-z_i)^2}
  +
  \frac{\partial_{z_i}\sigma(z_i)}{\zeta-z_i}.
\]
Fixing three points by \(PSL_2(\mathbb C)\) covariance reduces the resulting
Ward identity to \eqref{eq:ising-sigma-bpz-equation}.  Direct substitution
checks that \(f_+\) and \(f_-\) solve the second-order equation.

The two solutions are multivalued, but the full correlator must be
single-valued.  Around \(z=0\), the combinations
\(\mathcal F_0\) and \(\mathcal F_{\varepsilon}\) have the OPE behavior of
the identity and energy channels in
\(\sigma\times\sigma=\mathbf1\oplus\varepsilon\).  Therefore the most general
diagonal single-valued expression in this basis is
\[
  f(z,\bar z)=|\mathcal F_0(z)|^2+C^2|\mathcal F_\varepsilon(z)|^2
\]
with \(C^2\ge0\) by reflection positivity.  The holomorphic solutions obey
the reduced crossing transformation
\[
  f_+(1-z)
  =
  \left(\frac{1-z}{z}\right)^{1/8}
  \frac{f_+(z)+f_-(z)}{\sqrt2},
  \qquad
  f_-(1-z)
  =
  \left(\frac{1-z}{z}\right)^{1/8}
  \frac{f_+(z)-f_-(z)}{\sqrt2}.
\]
The scalar prefactor is not part of the fusion matrix: it is the conformal
weight factor coming from the external spin fields.  In the cross-ratio frame
\(G(z,\bar z)=|z|^{-1/4}f(z,\bar z)\), crossing says
\(|1-z|^{-1/4}f(1-z,1-\bar z)=|z|^{-1/4}f(z,\bar z)\), and this exactly
cancels the modulus square of \(((1-z)/z)^{1/8}\).  Thus the matrix part is
the invariant datum.
Equivalently,
\[
  \mathcal F_0(1-z)
  =
  \left(\frac{1-z}{z}\right)^{1/8}
  \left(
  \frac1{\sqrt2}\mathcal F_0(z)
  +\frac1{2\sqrt2}\mathcal F_\varepsilon(z)
  \right),
\]
\[
  \mathcal F_\varepsilon(1-z)
  =
  \left(\frac{1-z}{z}\right)^{1/8}
  \left(
  \sqrt2\,\mathcal F_0(z)
  -\frac1{\sqrt2}\mathcal F_\varepsilon(z)
  \right).
\]
The remaining crossing condition for the matrix part then forces the
coefficient of the mixed term
\(\mathcal F_0\overline{\mathcal F_\varepsilon}+\mathrm{c.c.}\) to vanish:
\[
  \frac12-2C^2=0.
\]
Thus \(C^2=1/4\).  The coefficients of
\(|\mathcal F_0|^2\) and \(|\mathcal F_\varepsilon|^2\) then also match,
giving \eqref{eq:ising-spin-four-point-structure-constant}.
\end{proof}

\section{Sewing and Higher-Genus Blocks}
\label{sec:voa-higher-genus-sewing}

Sewing is the operation that constructs higher-genus surfaces and higher-point
blocks from lower-genus pieces.  Let \(C_1\) and \(C_2\) have marked points
with coordinates \(z\) and \(w\).  Removing disks and imposing
\[
  zw=q,\qquad |q|<1,
\]
glues the two punctures into an annulus.  If the two glued punctures carry a
module \(M_a\) and its dual \(M_a^\vee\), a sewn chiral block has the form
\begin{equation}
  \mathcal F_{\rm sewn}(q)
  =
  \sum_{\alpha}
  q^{h_a+n_\alpha-c/24}
  \mathcal F_1(\cdots,e_\alpha)
  \mathcal F_2(e^\alpha,\cdots).
  \label{eq:chiral-block-sewing}
\end{equation}
Here \(\{e_\alpha\}\) is a homogeneous basis of \(M_a\), \(e^\alpha\) is the
dual basis with respect to the invariant pairing, and
\(L_0 e_\alpha=(h_a+n_\alpha)e_\alpha\).  The exponent records propagation
along the sewn annulus.

The formula is meaningful only after specifying a topology on the module and
a convergence domain for the sum.  For a unitary positive-energy chiral
module \(M_a\) write
\[
  M_a=\widehat{\bigoplus_{n\ge0}}\,M_{a,h_a+n},
  \qquad
  d_a(n)=\dim M_{a,h_a+n},
\]
where \(L_0\) has eigenvalue \(h_a+n\) on \(M_{a,h_a+n}\).  For
\(0<|q|<1\) define the annular propagation operator
\begin{equation}
  P_a(q)=q^{L_0-c/24},
  \qquad
  P_a(q)e=q^{h_a+n-c/24}e
  \quad(e\in M_{a,h_a+n}),
  \label{eq:annular-propagator}
\end{equation}
with a fixed branch of \(\log q\) on the sewing coordinate chart.  Its trace
norm is
\begin{equation}
  \|P_a(q)\|_1
  =
  |q|^{h_a-c/24}\sum_{n\ge0}d_a(n)|q|^n
  =
  |\chi_a(q)|,
  \label{eq:annular-propagator-trace-norm}
\end{equation}
up to the phase carried by \(q^{L_0-c/24}\).  Thus the first analytic
condition behind sewing is ordinary character convergence at the modulus
\(|q|\).

\begin{definition}[Hilbert sewing datum]
\label{def:hilbert-sewing-datum}
Let \(M_a\) be a unitary chiral module with nondegenerate invariant Hermitian
pairing.  A one-channel Hilbert sewing datum consists of:
\begin{enumerate}
  \item two continuous linear functionals
  \(\ell_1,\ell_2\) on the Hilbert completion of \(M_a\), representing the
  two lower-genus chiral blocks with one \(M_a\)-leg and one \(M_a^\vee\)-leg;
  \item a sewing parameter \(q\) for which the propagator \(P_a(q)\) is
  trace class;
  \item a choice of dual homogeneous bases
  \(\{e_\alpha\}\), \(\{e^\alpha\}\) satisfying
  \(\langle e^\alpha,e_\beta\rangle=\delta^\alpha{}_\beta\).
\end{enumerate}
The sewn block is the scalar
\begin{equation}
  \operatorname{Sew}_{a,q}(\ell_1,\ell_2)
  =
  \sum_\alpha
  \ell_1(e_\alpha)\,
  q^{L_0(e_\alpha)-c/24}\,
  \ell_2(e^\alpha),
  \label{eq:hilbert-sewing-scalar}
\end{equation}
provided the series is absolutely convergent.
\end{definition}

\begin{lemma}[Trace-class estimate for one-channel sewing]
\label{lem:trace-class-estimate-one-channel-sewing}
Assume \(P_a(q)\) is trace class and that
\(\ell_i(v)=\langle u_i,v\rangle\) for vectors \(u_i\) in the Hilbert
completion of \(M_a\).  Then the series
\eqref{eq:hilbert-sewing-scalar} is absolutely convergent and satisfies
\begin{equation}
  \left|
    \operatorname{Sew}_{a,q}(\ell_1,\ell_2)
  \right|
  \le
  \|u_1\|\,\|u_2\|\,\|P_a(q)\|_1 .
  \label{eq:one-channel-sewing-bound}
\end{equation}
In particular the sewn block is independent of the chosen orthonormal
homogeneous basis.
\end{lemma}

\begin{proof}
Choose the homogeneous basis orthonormal.  With
\(\lambda_\alpha=q^{L_0(e_\alpha)-c/24}\),
\[
  \operatorname{Sew}_{a,q}(\ell_1,\ell_2)
  =
  \sum_\alpha
  \langle u_1,e_\alpha\rangle\lambda_\alpha
  \langle e_\alpha,u_2\rangle
  =
  \langle u_1,P_a(q)u_2\rangle
\]
when the finite-level partial sums are taken first.  Trace class implies
compactness and boundedness, with
\(\|P_a(q)\|\le\|P_a(q)\|_1\).  Hence
\[
  |\langle u_1,P_a(q)u_2\rangle|
  \le
  \|u_1\|\,\|P_a(q)\|_1\,\|u_2\|.
\]
Absolute convergence of the displayed basis expansion follows directly from
\[
  \sum_\alpha
  |\langle u_1,e_\alpha\rangle|\,|\lambda_\alpha|\,
  |\langle e_\alpha,u_2\rangle|
  \le
  \|u_1\|\,\|u_2\|
  \sup_{\|v\|=\|w\|=1}
  \sum_\alpha |\lambda_\alpha|
  |\langle v,e_\alpha\rangle|\,|\langle e_\alpha,w\rangle|
  \le
  \|u_1\|\,\|u_2\|\sum_\alpha|\lambda_\alpha|.
\]
The last sum is the trace norm in this diagonal homogeneous basis.  If
\(P_a(q)\) is not diagonal in the chosen orthonormal basis, apply the same
argument to its singular-value decomposition
\[
  P_a(q)=\sum_r s_r\,|v_r\rangle\langle w_r|,
  \qquad
  s_r\ge0,\qquad
  \sum_r s_r=\|P_a(q)\|_1,
\]
which gives the same estimate by Cauchy--Schwarz in the \(r\)-sum.  The
finite-level partial sums converge in trace norm to \(P_a(q)\), so their
matrix elements converge to \(\langle u_1,P_a(q)u_2\rangle\).  Since this
inner product is an operator-theoretic scalar, not a coordinate expression,
changing the orthonormal homogeneous basis does not change the sewn block.
\end{proof}

\begin{remark}[Torus trace from self-sewing]
\label{rem:torus-trace-from-self-sewing}
Let \(M_a\) be a unitary positive-energy module and let \(v\) be a chiral
state whose vertex operator \(Y(v,1)\) preserves the common domain of
\(P_a(q)\).  Sewing the two punctures of the sphere two-point block
\(\langle e^\alpha,Y(v,1)e_\alpha\rangle\) gives
\begin{equation}
  Z_a(v;q)
  =
  \operatorname{Tr}_{M_a}
  \left(
    Y(v,1)q^{L_0-c/24}
  \right).
  \label{eq:torus-one-point-trace-from-sewing}
\end{equation}
For \(v=\mathbf 1\), this is the character
\begin{equation}
  Z_a(\mathbf 1;q)=\chi_a(q)
  =
  \operatorname{Tr}_{M_a}q^{L_0-c/24}.
  \label{eq:character-from-self-sewing}
\end{equation}
Indeed, if \(\{e_\alpha\}\) is a homogeneous orthonormal basis, the self-sewn
block is
\[
  \sum_\alpha
  q^{L_0(e_\alpha)-c/24}
  \langle e_\alpha,Y(v,1)e_\alpha\rangle,
\]
which is exactly the trace of
\(Y(v,1)q^{L_0-c/24}\) on finite-level truncations and hence, whenever the
trace is convergent, the operator trace.  Setting \(v=\mathbf 1\) gives
\(Y(\mathbf 1,1)=\operatorname{id}_{M_a}\).
\end{remark}

\begin{example}[Torus free-boson Green kernel and neutrality]
\label{ex:torus-free-boson-green-kernel}
The preceding trace formula is still abstract.  For the diagonal free boson one
can see the first period-kernel calculation explicitly.  Let
\[
  E_\tau=\mathbb C/(\mathbb Z+\tau\mathbb Z),\qquad
  \tau_2=\operatorname{Im}\tau>0,
\]
and use the normalization
\(X(z,\bar z)X(0)\sim-\log |z|^2\).  The scalar zero mode is not invertible, so
the Green kernel is defined on the orthogonal complement of constants.  A
convenient representative is
\begin{equation}
  G_\tau(z)
  =
  -\log\left|
    \frac{\vartheta_1(z\mid\tau)}{\eta(\tau)}
  \right|^2
  +\frac{2\pi(\operatorname{Im}z)^2}{\tau_2}
  +C_\tau ,
  \label{eq:torus-free-boson-green-kernel}
\end{equation}
where \(C_\tau\) fixes the normal-ordering convention, for instance by
\(\int_{E_\tau}G_\tau\,\dd^2z=0\).  With
\(\Delta=4\partial_z\partial_{\bar z}\) this kernel satisfies
\begin{equation}
  -\frac1{4\pi}\Delta_zG_\tau(z)
  =
  \delta_{E_\tau}(z)-\frac1{\tau_2}.
  \label{eq:torus-green-zero-mode-subtraction}
\end{equation}
The constant source term on the right side is not optional: it is the projection
away from the target-space zero mode.  Without it the Laplacian cannot be
inverted on a compact surface.

For normal-ordered vertex insertions
\[
  V_{a_i}(z_i,\bar z_i)=:\!\exp(\ii a_iX(z_i,\bar z_i))\!:
\]
the noncompact genus-one correlator, with the overall determinant and
target-volume factor suppressed, is
\begin{equation}
  \left\langle\prod_{i=1}^n V_{a_i}(z_i,\bar z_i)\right\rangle_\tau
  =
  \mathbf 1_{\sum_i a_i=0}
  \exp\!\left(
    -\sum_{i<j}a_ia_j\,G_\tau(z_i-z_j)
  \right),
  \label{eq:torus-free-boson-vertex-correlator}
\end{equation}
in the same normal-ordering convention.  The factor
\(\mathbf 1_{\sum_i a_i=0}\) is the integral over the constant target mode.  It
is the genus-one version of charge conservation in the operator product: a
non-neutral product is not made meaningful by adding more oscillator sewing
data.  The diagonal self-contractions have already been subtracted into the
definition of the normal-ordered fields; putting them back changes the
renormalized vertex normalization rather than computing a new correlator.

Formula \eqref{eq:torus-free-boson-green-kernel} is the genus-one shadow of the
higher-genus prime-form expression.  On a compact surface with period matrix
\(\Omega\), normalized holomorphic one-forms \(\omega_I\), and prime form
\(E(p,q)\), the corresponding scalar kernel has the period-kernel form
\[
  G_\Omega(p,q)
  =
  -\log |E(p,q)|^2
  +2\pi\,
   \operatorname{Im}\!\int_q^p\omega_I\,
   (\operatorname{Im}\Omega)^{-1}_{IJ}\,
   \operatorname{Im}\!\int_q^p\omega_J
  +C_\Omega .
\]
This displayed expression identifies the next analytic objects that sewing must
control: the prime form, period matrix, normalization of zero modes, and
normal-ordering constant.  For fermions one must also choose a spin structure
and replace the scalar Green kernel by the Szego kernel.  These are not
consequences of the finite graph-counting formula
\eqref{eq:sewing-graph-genus}; they are the concrete worldsheet kernels that a
higher-genus CFT calculation has to construct.
\end{example}

\begin{definition}[Sewing graph for a chiral block]
\label{def:chiral-sewing-graph}
A chiral sewing graph \(\Gamma\) consists of:
\begin{enumerate}
  \item vertices \(v\), each labelled by a genus \(g_v\), external insertion
  labels, and a finite-dimensional chiral block functional
  \(\mathcal F_v\);
  \item internal edges \(e\), each joining two marked local coordinates or
  two marked coordinates on the same vertex surface;
  \item for every internal edge \(e\), a channel label \(a_e\), a pair of
  dual chiral modules \(M_{a_e}\), \(M_{a_e}^\vee\), and a sewing parameter
  \(q_e\) with \(|q_e|<1\);
  \item an orientation convention specifying which end of each edge carries
  \(M_{a_e}\) and which carries \(M_{a_e}^\vee\).
\end{enumerate}
Its geometric genus is
\begin{equation}
  g(\Gamma)=
  \sum_v g_v+b_1(\Gamma),
  \label{eq:sewing-graph-genus}
\end{equation}
where \(b_1(\Gamma)\) is the first Betti number of the graph.
\end{definition}

Given \(\Gamma\), the formal sewn block is obtained by inserting a dual basis
on every internal edge and multiplying the propagator factors:
\begin{equation}
  \mathcal F_\Gamma(\{q_e\})
  =
  \sum_{\{\alpha_e\}}
  \left(
    \prod_{e}
    q_e^{L_0(e_{\alpha_e})-c/24}
  \right)
  \prod_v
  \mathcal F_v(\ldots,e_{\alpha_e},e^{\alpha_e},\ldots).
  \label{eq:sewing-graph-block}
\end{equation}
The genus formula \eqref{eq:sewing-graph-genus} follows because sewing two
connected components adds genera, while sewing two punctures on the same
connected component adds one independent cycle.

\paragraph{Finite-channel sewing estimate.}
Assume a rational unitary chiral theory with finitely many channel labels,
and fix a sewing graph \(\Gamma\).  Suppose that for each edge \(e\),
\(P_{a_e}(q_e)\) is trace class, and that every vertex block functional
\(\mathcal F_v\) extends continuously to the Hilbert tensor product of its
incident channel modules.  Then the edge-basis sum
\eqref{eq:sewing-graph-block} converges absolutely for the chosen
\(\{q_e\}\).  Its value is independent of the orthonormal homogeneous bases
used on internal edges.

Insert finite-level projectors \(E_{a_e,\le N}\) on every internal edge.  For
fixed cutoffs the sum is finite.  Choose orthonormal homogeneous bases on the
cutoff subspaces and write
\[
  \lambda_{e,\alpha}
  =
  q_e^{L_0(e_{\alpha})-c/24}
\]
for the diagonal matrix coefficient of the annular propagator on edge \(e\).
If \(\mathbf e_v(\boldsymbol\alpha)\) denotes the tensor product of the
unit basis vectors incident on \(v\), then continuity of the vertex block
functional gives the pointwise bound
\[
  \left|
    \mathcal F_v(\mathbf e_v(\boldsymbol\alpha))
  \right|
  \le
  \|\mathcal F_v\| .
\]
Therefore the absolute value of the finite cutoff contraction is bounded by
\begin{align}
  \sum_{\{\alpha_e\}}
  \left(
    \prod_e |\lambda_{e,\alpha_e}|
  \right)
  \left(
    \prod_v
    \left|
      \mathcal F_v(\mathbf e_v(\boldsymbol\alpha))
    \right|
  \right)
  &\le
  \left(\prod_v \|\mathcal F_v\|\right)
  \prod_e
  \left(\sum_{\alpha_e}|\lambda_{e,\alpha_e}|\right)       \notag\\
  &\le
  \left(\prod_v \|\mathcal F_v\|\right)
  \left(\prod_e \|P_{a_e}(q_e)\|_1\right).
  \label{eq:multi-edge-sewing-trace-bound}
\end{align}
The first inequality uses only that the same edge index is seen by the two
incident vertices; after taking absolute values the indices separate into a
product of one-edge sums.  The second inequality is equality in a homogeneous
diagonal basis and is bounded by the trace norm in any orthonormal basis.
Since the right hand side of
\eqref{eq:multi-edge-sewing-trace-bound} is finite and independent of the
cutoffs, the finite absolute sums are dominated by the full product of the
edge \(\ell^1\) sums.  Enlarging the cutoff on one or more edges adds a tail
whose absolute value is bounded by the same product with at least one edge
sum replaced by its trace-norm tail; these tails tend to zero.  Hence the
cutoff net is absolutely Cauchy.  Equivalently, if one replaces
\(P_{a_e}(q_e)\) by its singular-value truncations, the tensor product of the
edge operators converges in trace norm because
\[
  \left\|
    \bigotimes_{e} P_{a_e}(q_e)
  \right\|_1
  =
  \prod_e \|P_{a_e}(q_e)\|_1
\]
for finite tensor products of trace-class operators, and the vertex
functionals are bounded on the Hilbert tensor products.  This operator
description also proves basis independence: two homogeneous orthonormal
bases give the same trace-class tensor contraction after the cutoff limit.

The estimate is intentionally local.  It controls a plumbing chart with
fixed channel labels and \(|q_e|<1\).  It does not prove that the resulting
functions analytically continue between different pants decompositions, nor
that two such continuations agree on the overlap of coordinate charts.  Those
are the analytic sewing and Moore--Seiberg compatibility statements isolated
next.

\begin{hypothesis}[Analytic sewing theorem for rational chiral data]
\label{hyp:rational-analytic-sewing-theorem}
For the rational chiral theories treated as fully constructed in this
chapter, the absolutely convergent local sewing series
\eqref{eq:sewing-graph-block} analytically continue over the appropriate
cover of the moduli space of pointed curves, assemble into finite-rank vector
bundles of chiral blocks, and are compatible with changes of pants
decomposition by the fusing, braiding, and modular moves.  The compatibility
maps obey the pentagon, hexagon, and genus-one modular identities.
\end{hypothesis}

This hypothesis is not a cosmetic assumption.  The trace-class estimates
above control a local plumbing coordinate chart near \(q_e=0\).  They do not
by themselves prove global analytic continuation, mapping-class-group
covariance, or equality of two different decompositions of the same Riemann
surface.  Those assertions are precisely the higher-genus sewing theorem.

\begin{definition}[Sewing-compatible rational chiral theory]
\label{def:sewing-compatible-rational-chiral-theory}
For this chapter, a rational chiral theory is a VOA \(V\) with finitely many
simple ordinary modules \(\{M_i\}_{i\in I}\), semisimple module category,
finite-dimensional conformal-block spaces, nondegenerate two-point pairings,
and convergent sewing expansions \eqref{eq:chiral-block-sewing} in a
neighborhood of \(q=0\), satisfying
Hypothesis~\ref{hyp:rational-analytic-sewing-theorem}.  The analytic
continuations of these blocks along the moduli space of pointed curves give a
projectively flat vector bundle of conformal blocks.
\end{definition}

The mapping class group acts on this bundle.  In genus zero the elementary
moves are associativity and braiding of OPE channels.  In genus one the
generators are
\[
  S:\tau\mapsto -1/\tau,\qquad
  T:\tau\mapsto\tau+1.
\]
The compatibility of sewing with these moves is the Moore--Seiberg system:
pentagon, hexagon, and modular identities for associativity, braiding, and
twist data.  The modern algebraic home of this system is a modular tensor
category when the module category is semisimple.

\subsection{Moore--Seiberg Data as Defect-Junction Consistency}
\label{subsec:moore-seiberg-defect-junction-consistency}

The Moore--Seiberg equations should be read as equalities of conformal-block
transport maps, or equivalently as equalities of topological-line junction
networks after the physical null quotient of
Definition~\ref{def:physical-defect-category-null-quotient}.  This is the
same physical-defect category used in
Chapter~\ref{ch:global-categorical-symmetry-defect-fusion}; here its objects
are the topological lines labelled by chiral sectors.

Let \(\mathcal C\) be a semisimple chiral sector category with simple objects
\(\{M_i\}_{i\in I}\).  For simple labels define the fusion spaces
\begin{equation}
  V_{ij}^{k}:=\operatorname{Hom}_{\mathcal C}(M_i\boxtimes M_j,M_k),
  \qquad
  N_{ij}{}^k=\dim V_{ij}^k .
  \label{eq:moore-seiberg-fusion-spaces}
\end{equation}
A vector in \(V_{ij}^{k}\) is a trivalent junction at which the two lines
\(i,j\) fuse to the line \(k\).  For four external labels \(i,j,k,\ell\),
the two standard pants decompositions have block spaces
\begin{equation}
  \bigoplus_m V_{ij}^{m}\otimes V_{m k}^{\ell},
  \qquad
  \bigoplus_n V_{jk}^{n}\otimes V_{i n}^{\ell}.
  \label{eq:moore-seiberg-two-four-point-bases}
\end{equation}
The fusing matrix is the change of basis
\begin{equation}
  F_{ijk}^{\ell}:
  \bigoplus_m V_{ij}^{m}\otimes V_{m k}^{\ell}
  \longrightarrow
  \bigoplus_n V_{jk}^{n}\otimes V_{i n}^{\ell}.
  \label{eq:moore-seiberg-f-move}
\end{equation}
The braiding maps are linear isomorphisms
\begin{equation}
  R_{ij}^{k}:V_{ij}^{k}\longrightarrow V_{ji}^{k},
  \label{eq:moore-seiberg-r-move}
\end{equation}
and the twist is a scalar \(\theta_i\) on each simple line in the semisimple
case.  They are the local coordinates of a topological defect category:
\(F\) changes the parenthesization of three fusions, \(R\) exchanges two
adjacent lines, and
\(\theta\) records a full rotation of one line.

The pentagon is the assertion that the two composites from
\((((a\boxtimes b)\boxtimes c)\boxtimes d)\) to
\(a\boxtimes(b\boxtimes(c\boxtimes d))\) agree:
\begin{equation}
  \alpha_{a,b,c\boxtimes d}\circ\alpha_{a\boxtimes b,c,d}
  =
  (\mathbf 1_a\boxtimes\alpha_{b,c,d})\circ
  \alpha_{a,b\boxtimes c,d}\circ
  (\alpha_{a,b,c}\boxtimes\mathbf 1_d).
  \label{eq:moore-seiberg-pentagon-associator}
\end{equation}
Here \(\alpha_{a,b,c}:(a\boxtimes b)\boxtimes c\to
a\boxtimes(b\boxtimes c)\) is represented, in the bases
\eqref{eq:moore-seiberg-two-four-point-bases}, by the corresponding
\(F\)-matrix.  In multiplicity sectors the pentagon is an equality of linear
maps between direct sums of junction spaces.  Physically it says that the
correlator obtained by fusing four nearby topological lines is independent of
the order in which the intermediate collars are shrunk.

The two hexagons say that moving one line past a fused pair is the same as
moving it past the two constituents one at a time, with the associators
included.  With braiding
\(c_{a,b}:a\boxtimes b\to b\boxtimes a\), one of the two hexagons is
\begin{equation}
  c_{a,b\boxtimes c}
  =
  \alpha_{b,c,a}^{-1}\circ
  (\mathbf 1_b\boxtimes c_{a,c})\circ
  \alpha_{b,a,c}\circ
  (c_{a,b}\boxtimes \mathbf 1_c)\circ
  \alpha_{a,b,c}^{-1},
  \label{eq:moore-seiberg-hexagon-one}
\end{equation}
as a map \(a\boxtimes(b\boxtimes c)\to (b\boxtimes c)\boxtimes a\).
The other hexagon is the parallel identity for
\(c_{a\boxtimes b,c}\).  The balancing identity
\[
  \theta_{a\boxtimes b}
  =
  c_{b,a}\circ c_{a,b}\circ(\theta_a\boxtimes\theta_b)
\]
then fixes how a Dehn twist acts on a fused pair.  In a rational unitary CFT,
these equalities are the local algebraic part of sewing: they identify
different cuts of the same punctured surface after analytic continuation
around the moduli space.

This formulation also fixes the status of the algebraic classification
problem.  A solution of the pentagon, hexagon, and modular equations is a
candidate topological-line and conformal-block datum.  To obtain a QFT one
must still construct the Hilbert space of states, local fields or local
algebras, positive pairings, convergent correlation functions, and sewing maps
whose analytic continuations realize the displayed \(F\), \(R\), and
\(\theta\).  Conversely, once those analytic and positivity structures are
constructed, the Moore--Seiberg equations follow from equality of two
correlation-function limits in the defect category.  The monograph therefore
uses the equations as structural constraints and as coordinates on known
classes of rational CFT data; the definition of the CFT itself includes the
analytic and Hilbert-space structures just listed.

\section{Modular Tensor Category and Verlinde Formula}

Let \(\mathcal C\) be the semisimple ribbon category of modules of a rational
chiral theory.  Its simple objects are \(M_i\), with tensor product
\[
  M_i\boxtimes M_j\cong \bigoplus_k N_{ij}{}^k M_k .
\]
The matrices \((N_i)_j{}^k=N_{ij}{}^k\) are the fusion matrices.  The category
has a braiding, duals, and twists \(\theta_i=\ee^{2\pi\ii h_i}\).  Its
modular \(S\)-matrix is the Hopf-link pairing normalized so that the vacuum
label \(0\) corresponds to \(V\).  In a unitary theory \(S\) is unitary and
\[
  d_i:=\frac{S_{i0}}{S_{00}}
\]
is the quantum dimension of \(M_i\).

\begin{quotedtheorem}[Modularity of rational chiral characters]
\label{qthm:voa-character-modularity}
For a rational, \(C_2\)-cofinite, positive-energy VOA satisfying the analytic
finiteness hypotheses needed for sewing, the characters
\(\chi_i(\tau)\) of simple modules span a finite-dimensional representation
of \(SL(2,\mathbb Z)\):
\[
  \chi_i(-1/\tau)=\sum_j S_{ij}\chi_j(\tau),
  \qquad
  \chi_i(\tau+1)=\sum_j T_{ij}\chi_j(\tau).
\]
The representation agrees with the modular data of the semisimple module
category.  This is Zhu's modular-invariance theorem, with the
tensor-category identification supplied by the Huang--Lepowsky theory in the
rational setting \cite{Zhu1996,Huang2005}.
\end{quotedtheorem}

\paragraph{Finite trace-recursion mechanism.}
The proof boundary in Quoted Theorem~\ref{qthm:voa-character-modularity} is
analytic, but the algebraic source of finite modular closure is explicit.
Use Zhu's square-bracket coordinate
\begin{equation}
  Y[a,z]:=Y(\ee^{zL_0}a,\ee^z-1)
  =
  \sum_{n\in\mathbb Z}a[n]z^{-n-1},
  \label{eq:zhu-square-bracket-vertex-operator}
\end{equation}
and write
\[
  Z_i(v,\tau)=
  \operatorname{Tr}_{M_i}o(v)\,q^{L_0-c/24}.
\]
Let
\begin{equation}
  \wp(z,\tau)
  =
  z^{-2}
  +\sum_{\ell\ge1}(2\ell+1)G_{2\ell+2}(\tau)z^{2\ell},
  \qquad
  G_{2k}(\tau)=
  \sum_{(m,n)\in\mathbb Z^2\setminus\{0\}}(m\tau+n)^{-2k}.
  \label{eq:zhu-weierstrass-kernel}
\end{equation}
The genus-one Ward identity obtained by moving an \(a\)-contour around the
torus gives, for homogeneous \(a,b\in V\),
\begin{equation}
  Z_i(a[-2]b,\tau)
  +
  \sum_{\ell\ge1}(2\ell+1)G_{2\ell+2}(\tau)
  Z_i(a[2\ell]b,\tau)
  =
  0.
  \label{eq:zhu-genus-one-recursion}
\end{equation}
Equation~\eqref{eq:zhu-genus-one-recursion} is the torus analogue of the
sphere residue calculation in Proposition~\ref{prop:voa-ope-from-locality}:
the pole \(z^{-2}\) records the \(a[-2]b\) descendant, while the holomorphic
terms of \(\wp\) record the allowed windings of the contour around the
elliptic curve.

Now \(C_2\)-cofiniteness supplies a finite spanning set for the square-bracket
descendant quotient.  Repeated use of
\eqref{eq:zhu-genus-one-recursion}, together with the corresponding
\(L[-2]\)-insertion identity that gives the modular covariant
\(\tau\)-derivative, reduces all trace functions to a finite system of
linear differential equations with modular-form coefficients.  The quoted
theorem adds the hard analytic conclusions: convergence of the trace
functions on the upper half-plane, continuation to a finite-dimensional
space of holomorphic functions, and invariance of that space under the
generators \(S:\tau\mapsto-1/\tau\) and \(T:\tau\mapsto\tau+1\).  The
Verlinde calculation below uses only the finite modular representation after
these analytic inputs have been secured.

\paragraph{Verlinde formula from modular diagonalization.}
Assume the fusion matrices are simultaneously diagonalized by a unitary
matrix \(S\) with \(S_{0m}\ne0\):
\begin{equation}
  (N_i S)_{jm}
  =
  \frac{S_{im}}{S_{0m}}S_{jm}.
  \label{eq:fusion-diagonalization-by-s}
\end{equation}
Then
\begin{equation}
  N_{ij}{}^k
  =
  \sum_m
  \frac{S_{im}S_{jm}\overline{S_{km}}}{S_{0m}} .
  \label{eq:verlinde-formula}
\end{equation}

The coefficient \(N_{ij}{}^k\) is the \(j,k\) matrix element of \(N_i\):
\[
  N_{ij}{}^k=(N_i)_j{}^k.
\]
Equation~\eqref{eq:fusion-diagonalization-by-s} says
\[
  N_i
  =
  S D_i S^{-1},
  \qquad
  (D_i)_{mm}=\frac{S_{im}}{S_{0m}}.
\]
Since \(S\) is unitary, \(S^{-1}_{mk}=\overline{S_{km}}\).  Therefore
\[
  (N_i)_j{}^k
  =
  \sum_m S_{jm}\frac{S_{im}}{S_{0m}}\overline{S_{km}},
\]
which is \eqref{eq:verlinde-formula}.

This calculation separates the deep input from the algebra.  The deep input is
that the same \(S\)-matrix obtained by the modular transformation of torus
characters diagonalizes the fusion ring.  Once this is granted, the formula is
linear algebra.

\section{The Ising Chiral Theory}
\label{sec:ising-chiral-theory}

The Ising minimal model has central charge
\[
  c=\frac12
\]
and three simple modules:
\[
  \mathbf 1,\qquad \sigma,\qquad \varepsilon ,
\]
with conformal weights
\[
  h_{\mathbf 1}=0,\qquad h_\sigma=\frac1{16},\qquad
  h_\varepsilon=\frac12 .
\]
In the ordered basis \((\mathbf 1,\sigma,\varepsilon)\), the modular
\(S\)-matrix is
\begin{equation}
  S_{\rm Ising}
  =
  \frac12
  \begin{pmatrix}
    1 & \sqrt2 & 1\\
    \sqrt2 & 0 & -\sqrt2\\
    1 & -\sqrt2 & 1
  \end{pmatrix}.
  \label{eq:ising-modular-s}
\end{equation}
The corresponding \(T\)-matrix is diagonal:
\begin{equation}
  T_{\rm Ising}
  =
  \operatorname{diag}\left(
    \ee^{-\pi\ii/24},
    \ee^{\pi\ii/12},
    \ee^{23\pi\ii/24}
  \right),
  \label{eq:ising-modular-t}
\end{equation}
because \(T_{ii}=\ee^{2\pi\ii(h_i-c/24)}\) and \(c=1/2\).
It is real, unitary, and obeys \(S^2=1\), because every simple object is
self-dual.  The quantum dimensions are
\[
  d_{\mathbf 1}=1,\qquad d_\sigma=\sqrt2,\qquad d_\varepsilon=1,
  \qquad
  \mathcal D^2=\sum_i d_i^2=4.
\]
Applying \eqref{eq:verlinde-formula} gives the fusion rules
\[
  \varepsilon\boxtimes\varepsilon=\mathbf 1,\qquad
  \varepsilon\boxtimes\sigma=\sigma,\qquad
  \sigma\boxtimes\sigma=\mathbf 1\oplus\varepsilon .
\]
The verification is pure Verlinde algebra.  The identity \(S^2=1\) is a direct
matrix multiplication using
\((\sqrt2)^2=2\).  The vacuum row gives the quantum dimensions by
\(d_i=S_{i0}/S_{00}\).  For example,
\[
  N_{\sigma\sigma}{}^{\mathbf 1}
  =
  \sum_m\frac{S_{\sigma m}S_{\sigma m}S_{\mathbf 1m}}{S_{\mathbf 1m}}
  =
  S_{\sigma\mathbf 1}^2+S_{\sigma\sigma}^2+S_{\sigma\varepsilon}^2
  =
  \frac12+0+\frac12=1,
\]
and
\[
  N_{\sigma\sigma}{}^\sigma
  =
  \sum_m\frac{S_{\sigma m}S_{\sigma m}S_{\sigma m}}{S_{\mathbf 1m}}
  =
  \frac{\sqrt2}{2}-\frac{\sqrt2}{2}=0.
\]
The same substitution gives
\(N_{\sigma\sigma}{}^\varepsilon=1\),
\(N_{\varepsilon\varepsilon}{}^{\mathbf 1}=1\), and
\(N_{\varepsilon\sigma}{}^\sigma=1\), with all other coefficients fixed by
commutativity and the unit object.

\section{Operator-Algebraic Conformal Nets}
\label{sec:operator-algebraic-conformal-nets}

The VOA formalism starts from chiral fields as formal distributions and then
extracts algebraic consequences.  The operator-algebraic formalism starts
from bounded observables localized in intervals on the circle.  This
difference has operator-algebraic content.  Smeared chiral fields are usually unbounded
operators; commutation on a common invariant domain does not by itself imply
commutation of the generated von Neumann algebras.  A conformal net records
the bounded local algebras directly, and treats fields as affiliated
coordinates when such coordinates exist.

\begin{definition}[Chiral conformal net]
\label{def:chiral-conformal-net}
Let \(\mathcal I(S^1)\) be the set of connected, nonempty, nondense open
intervals \(I\subset S^1\).  A M\"obius covariant chiral local net consists
of a Hilbert space \(\Hilb\), a unit vector \(\Omega\in\Hilb\), a positive
energy unitary representation \(U\) of the M\"obius group on \(\Hilb\), and
von Neumann algebras
\[
  \mathcal A(I)\subset B(\Hilb),\qquad I\in\mathcal I(S^1),
\]
such that:
\begin{enumerate}
\item isotony:
  \(I_1\subset I_2\) implies \(\mathcal A(I_1)\subset\mathcal A(I_2)\);
\item locality:
  \(I_1\cap I_2=\varnothing\) implies
  \([\mathcal A(I_1),\mathcal A(I_2)]=0\);
\item covariance:
  \(U(\gamma)\mathcal A(I)U(\gamma)^*=\mathcal A(\gamma I)\);
\item positive energy:
  the generator of rotations has spectrum bounded below by \(0\);
\item vacuum:
  \(U(\gamma)\Omega=\Omega\), and
  \(\Omega\) is cyclic for
  \(\bigvee_{I\in\mathcal I(S^1)}\mathcal A(I)\).
\end{enumerate}
The net is irreducible if
\(\bigvee_I\mathcal A(I)=B(\Hilb)\).  It is a conformal net, or
diffeomorphism-covariant net, if \(U\) extends to a strongly continuous
projective positive-energy representation of
\(\operatorname{Diff}^+(S^1)\) satisfying the same covariance and the local
implementer condition: if a diffeomorphism is the identity on \(I\), its
implementer acts trivially on \(\mathcal A(I)\).
\end{definition}

The complement \(I'\) of an interval is the interior of
\(S^1\setminus I\).  Haag duality is the equality
\[
  \mathcal A(I)'=\mathcal A(I') .
\]
For irreducible M\"obius covariant nets on the circle it follows under the
standard vacuum-net hypotheses; in higher-dimensional AQFT it is a separate
condition, as in
Chapter~\ref{ch:volume-iv-superselection-sectors-locality-properties}.

\begin{definition}[Split property, strong additivity, and complete rationality]
\label{def:complete-rational-conformal-net}
Let \(\mathcal A\) be an irreducible M\"obius covariant net.
\begin{enumerate}
\item The split property means that whenever
  \(\overline I_1\subset I_2\) there is a type I factor \(N\) with
  \[
    \mathcal A(I_1)\subset N\subset\mathcal A(I_2).
  \]
\item Strong additivity means that if deleting one point from an interval
  \(I\) gives the two intervals \(I_1,I_2\), then
  \[
    \mathcal A(I)=\mathcal A(I_1)\vee\mathcal A(I_2).
  \]
\item Assume the split property.  Choose four intervals
  \(I_1,I_2,I_3,I_4\) in cyclic order and set \(E=I_1\cup I_3\).  The
  \(\mu\)-index is the Jones index
  \[
    \mu_{\mathcal A}=[\mathcal A(E')':\mathcal A(E)]
  \]
  when this index is finite and independent of the chosen four intervals.
\end{enumerate}
The net is completely rational if it has the split property, is strongly
additive, and has finite \(\mu_{\mathcal A}\).
\end{definition}

\begin{quotedtheorem}[Representation category of a completely rational net]
\label{qthm:complete-rational-net-mtc}
For a completely rational chiral conformal net, the finite-index localized
representations form a unitary modular tensor category.  The categorical
global dimension agrees with the \(\mu\)-index:
\[
  \mu_{\mathcal A}=\sum_{\rho\in\operatorname{Irr}(\mathcal A)} d_\rho^2,
\]
where \(d_\rho\) is the statistical dimension of the sector \(\rho\).
This is the Kawahigashi--Longo--M\"uger conformal-net rationality theorem.
The proof uses subfactor index theory and DHR localization, and is not
reproduced here \cite{KawahigashiLongoMuger2001}.
\end{quotedtheorem}

\paragraph{DHR mechanism behind the conformal-net theorem.}
The quoted theorem is not only a finiteness statement; it identifies the
operator-algebraic origin of the braiding and of the modular \(S\)-matrix.
Fix an interval \(I_0\).  A DHR sector localized in \(I_0\) is represented,
after Haag duality, by a normal unital endomorphism \(\rho\) of the
quasilocal algebra such that
\begin{equation}
  \rho(A)=A,\qquad A\in\mathcal A(I_0'),
  \label{eq:dhr-localized-endomorphism}
\end{equation}
where \(I_0'\) is the interior of the complement.  If \(\rho,\sigma\) are
two such endomorphisms, the morphism space is
\begin{equation}
  \operatorname{Hom}(\rho,\sigma)
  =
  \{T:T\rho(A)=\sigma(A)T\ {\rm for\ all\ local}\ A\}.
  \label{eq:dhr-intertwiner-space}
\end{equation}
The tensor product is composition of endomorphisms,
\(\rho\otimes\sigma:=\rho\circ\sigma\), with tensoring of intertwiners
defined by \(S\otimes T=S\rho(T)\) for
\(S\in\operatorname{Hom}(\rho,\rho')\) and
\(T\in\operatorname{Hom}(\sigma,\sigma')\).  Finite Jones index of the
subfactor \(\rho(\mathcal A(I))\subset\mathcal A(I)\), for an interval
\(I\supset I_0\), gives the statistical dimension
\begin{equation}
  d_\rho
  =
  [\mathcal A(I):\rho(\mathcal A(I))]^{1/2},
  \label{eq:dhr-statistical-dimension-index}
\end{equation}
independent of \(I\) in the finite-index sector.  Conjugate sectors are
endomorphisms \(\bar\rho\) equipped with intertwiners
\[
  R\in\operatorname{Hom}(\operatorname{id},\bar\rho\rho),
  \qquad
  \bar R\in\operatorname{Hom}(\operatorname{id},\rho\bar\rho),
\]
obeying the conjugate equations
\begin{equation}
  \bar R^*\rho(R)=\operatorname{id}_\rho,
  \qquad
  R^*\bar\rho(\bar R)=\operatorname{id}_{\bar\rho},
  \label{eq:dhr-conjugate-equations}
\end{equation}
after the conventional normalization that minimizes
\(\|R\|\,\|\bar R\|\).  This is the \(C^*\)-tensor-categorical form of
particle-antiparticle pairing in the chiral net.

The braiding is constructed from charge transporters.  Move \(\rho\) and
\(\sigma\), by unitary intertwiners, to disjoint intervals with a specified
cyclic order on \(S^1\).  Locality says that the transported
endomorphisms commute on all local algebras.  Transporting back gives a
unitary intertwiner
\begin{equation}
  \varepsilon_{\rho,\sigma}
  \in
  \operatorname{Hom}(\rho\sigma,\sigma\rho).
  \label{eq:dhr-braiding-intertwiner}
\end{equation}
Changing the transporters changes the intermediate representative but not
the resulting natural braiding morphism.  Naturality and the hexagon
identities are therefore consequences of isotony, locality, and the
composition law for charge transporters.  The two possible cyclic orders on
the circle give the braiding and its opposite.

Complete rationality supplies the missing finiteness and nondegeneracy.
For two disjoint intervals \(E=I_1\cup I_3\), the finite index
\[
  \mu_{\mathcal A}=[\mathcal A(E')':\mathcal A(E)]
\]
is the two-interval index measuring the failure of additivity across the two
gaps.  The subfactor analysis shows that only finitely many irreducible
finite-index DHR sectors occur and that
\begin{equation}
  \mu_{\mathcal A}
  =
  \sum_{\rho\in\operatorname{Irr}(\mathcal A)}d_\rho^2 .
  \label{eq:dhr-mu-index-global-dimension}
\end{equation}
The categorical \(S\)-matrix is the normalized trace of monodromy:
\begin{equation}
  S_{\rho\sigma}
  =
  \frac{1}{\mathcal D}\,
  \operatorname{Tr}_{\rho\sigma}
  \bigl(\varepsilon_{\sigma,\rho}\varepsilon_{\rho,\sigma}\bigr),
  \qquad
  \mathcal D=\sqrt{\sum_\lambda d_\lambda^2}.
  \label{eq:dhr-categorical-s-matrix}
\end{equation}
Here \(\operatorname{Tr}_{\rho\sigma}\) is the standard categorical trace
coming from the conjugate equations.  The vacuum row is
\(S_{\operatorname{id},\rho}=d_\rho/\mathcal D\).  Nondegeneracy means that
the only sector whose double braiding with every sector is the identity is
the vacuum sector; equivalently the matrix
\((S_{\rho\sigma})\) is invertible.  This is the precise operator-algebraic
content of the word ``modular'' in the theorem.  The full proof that the
two-interval index forces this nondegeneracy remains the subfactor-theoretic
theorem boundary, but the objects entering the statement are the concrete
endomorphisms, intertwiners, dimensions, and charge-transporter braidings
defined above.

The comparison with VOAs is therefore a comparison of two different
coordinates on chiral locality.  A VOA gives formal fields and mode algebra;
a net gives bounded interval algebras.  Passing from one to the other is an
analytic theorem, not a definition.

\begin{quotedtheorem}[Strongly local VOAs and conformal nets]
\label{qthm:voa-to-conformal-net}
Let \(V\) be a simple unitary VOA whose vertex operators obey polynomial
energy bounds.  If the smeared fields \(Y(a,f)\), \(a\in V\),
\(f\in C^\infty(S^1)\), generate mutually local von Neumann algebras
\[
  \mathcal A_V(I)
  =
  W^*\{\overline{Y(a,f)}:\operatorname{supp}f\subset I,\ a\in V\},
  \qquad I\in\mathcal I(S^1),
\]
then \(V\) is called strongly local and \(I\mapsto\mathcal A_V(I)\) is an
irreducible conformal net.  The net is independent, up to isomorphism, of
the normalized invariant scalar product on \(V\).

Conversely, the Fredenhagen--J\"or{\ss} reconstruction applied to
\(\mathcal A_V\) recovers the smeared vertex operators of \(V\).  More
generally, if an irreducible conformal net is generated by quasi-primary
reconstructed fields, those fields satisfy polynomial energy bounds, and the
\(L_0\)-eigenspaces are finite-dimensional, then the finite-energy subspace
\(\Hilb^{\mathrm{fin}}\) carries a simple unitary strongly local VOA
structure \(V\) with \(\mathcal A\simeq\mathcal A_V\).

This is the Carpi--Kawahigashi--Longo--Weiner VOA-to-net theorem and its
Fredenhagen--J\"or{\ss} converse under the stated analytic hypotheses
\cite{CarpiKawahigashiLongoWeiner2018}.
\end{quotedtheorem}

\paragraph{Mechanism of the VOA-to-net passage.}
Let \(\Hilb\) be the Hilbert completion of the unitary VOA \(V\), let
\(L_0\) be the positive rotation generator, and set
\[
  \Hilb^\infty:=\bigcap_{k\ge0}\operatorname{Dom}(1+L_0)^k .
\]
A polynomial energy bound for \(a\in V\) means that there exist constants
\(C,r,s\), depending on \(a\), such that for every mode \(a_n\) and every
\(\psi\in\Hilb^\infty\),
\begin{equation}
  \|a_n\psi\|
  \le
  C(1+|n|)^r\|(1+L_0)^s\psi\|.
  \label{eq:voa-polynomial-energy-bound}
\end{equation}
For a smooth test function
\[
  f(\ee^{\ii\theta})=\sum_{n\in\mathbb Z}\hat f_n\ee^{-\ii n\theta}
\]
the rapid decay of \(\hat f_n\) implies
\begin{equation}
  Y(a,f)\psi
  :=
  \sum_{n\in\mathbb Z}\hat f_n a_n\psi
  \label{eq:voa-smeared-field-series}
\end{equation}
converges in \(\Hilb\), because
\(\sum_n|\hat f_n|(1+|n|)^r<\infty\).  Unitarity supplies the adjoint
formula for vertex operators, so \(Y(a,f)\) is closable on
\(\Hilb^\infty\).  The local von Neumann algebra in the theorem is therefore
not obtained by treating \(Y(a,z)\) as a pointwise operator; it is generated
from closed smeared operators with supports contained in \(I\):
\[
  \mathcal A_V(I)
  =
  W^*\{\overline{Y(a,f)}:\operatorname{supp}f\subset I,\ a\in V\}.
\]

The delicate step is locality.  VOA locality gives, for each pair \(a,b\),
an integer \(N\) such that
\[
  (z-w)^N[Y(a,z),Y(b,w)]=0
\]
as a formal distribution identity.  After smearing with disjointly supported
smooth functions, this gives
\[
  [Y(a,f),Y(b,g)]\psi=0,\qquad \psi\in\Hilb^\infty .
\]
This common-domain commutator identity is not yet locality of the net: the
latter requires the von Neumann algebras generated by the closures to
commute.  The analytic content of strong locality is precisely the upgrade
from finite-energy-domain locality of unbounded fields to strong
commutativity of the closed smeared fields and their bounded functional
calculi.  Polynomial energy bounds, positivity of \(L_0\), and the adjoint
formula give the operator control used in that upgrade.

Covariance is more formal once the smeared fields exist.  The projective
unitary M\"obius and diffeomorphism action on the Hilbert completion sends a
smeared field to a finite linear combination of smeared fields whose test
functions are pulled back with the appropriate conformal-weight factors.
Thus
\[
  U(\gamma)\mathcal A_V(I)U(\gamma)^*=\mathcal A_V(\gamma I)
\]
provided the implementers preserve the common smooth domain and the energy
bounds.  Vacuum cyclicity follows from the state-field correspondence and
the density of VOA descendants in the Hilbert completion.  The theorem
boundary is the simultaneous proof that these analytic requirements are
stable for the whole field family, not only for a chosen finite generating
set.

\paragraph{Mechanism of the net-to-VOA reconstruction.}
The converse starts with bounded interval algebras rather than formal
fields.  For a finite-energy vector \(\psi\in\Hilb^{\mathrm{fin}}\), the
Fredenhagen--J\"or{\ss} construction seeks a closed operator-valued
distribution \(\Phi_\psi\) affiliated with the local algebras and normalized
by
\[
  \Phi_\psi(f)\Omega
  =
  \int_{S^1}f(z)\,U(z)\psi\,dz
\]
in the distributional sense, where \(U(z)\) denotes rotation of the state.
Locality of the net gives commutation of \(\Phi_\psi(f)\) with
\(\mathcal A(I')\) when \(\operatorname{supp}f\subset I\).  M\"obius
covariance decomposes \(\Hilb^{\mathrm{fin}}\) into finite-dimensional
\(L_0\)-eigenspaces and fixes the mode degrees of the reconstructed fields.
If the reconstructed quasi-primary fields obey polynomial energy bounds,
their smeared products have vacuum matrix elements with the same
analytic-continuation and locality properties as VOA correlation functions.
The modes of these fields then define a state-field map on
\(\Hilb^{\mathrm{fin}}\), with the vacuum, translation, locality, and
Virasoro covariance axioms recovered from the net axioms plus the
Fredenhagen--J\"or{\ss} field-state normalization.

The converse theorem boundary is again analytic rather than formal.  One
must prove that enough finite-energy vectors admit such reconstructed fields,
that the fields satisfy the required energy bounds, and that the mode
products close on \(\Hilb^{\mathrm{fin}}\).  Without those hypotheses a
conformal net need not be identified with a VOA merely because both describe
chiral locality.

\paragraph{Ising net \(\mu\)-index and DHR fusion.}
Assume the unitary Ising VOA is taken in its strongly local realization and
that its associated conformal net is completely rational with sectors
\(\mathbf 1,\sigma,\varepsilon\).  Then
\[
  \mu_{\mathcal A_{\mathrm{Ising}}}=4,
\]
and the DHR fusion ring of the net agrees with the Ising Verlinde fusion
ring displayed above.

From the explicit Ising \(S\)-matrix \eqref{eq:ising-modular-s}, the quantum
dimensions are
\[
  d_{\mathbf 1}=1,\qquad d_\sigma=\sqrt2,\qquad d_\varepsilon=1.
\]
The quoted complete-rationality theorem gives
\[
  \mu_{\mathcal A_{\mathrm{Ising}}}
  =
  d_{\mathbf 1}^2+d_\sigma^2+d_\varepsilon^2
  =
  1+2+1=4.
\]
The same theorem identifies finite-index DHR sectors with the modular tensor
category of the net.  Under the VOA-to-net comparison theorem, the strongly
local Ising VOA and the Ising net have the same chiral sectors, so the DHR
fusion coefficients are the Verlinde coefficients already computed.

\begin{openproblem}[Unqualified VOA/net equivalence]
\label{op:voa-net-equivalence-without-extra-hypotheses}
Remove or sharply weaken the analytic hypotheses in
Theorem~\ref{qthm:voa-to-conformal-net}.  In particular, determine whether
every simple unitary positive-energy VOA satisfying the specified finiteness
conditions is strongly local, and whether every irreducible conformal net of
finite chiral type is recovered from a strongly local VOA.  The monograph
uses the proved strongly local correspondence as a theorem boundary and does
not assume this unqualified equivalence.
\end{openproblem}

\section{Full CFT and the Modular Bootstrap}

A full local CFT is not a chiral theory alone.  In a rational diagonal case,
one constructs a torus partition function from left and right characters:
\begin{equation}
  Z(\tau,\bar\tau)
  =
  \sum_{i,j}M_{ij}\chi_i(\tau)\overline{\chi_j(\tau)} .
  \label{eq:rational-torus-partition-function}
\end{equation}
The matrix \(M\) must have nonnegative integer entries, \(M_{00}=1\), and must
commute with the modular representation:
\begin{equation}
  MS=SM,\qquad MT=TM .
  \label{eq:modular-invariant-matrix}
\end{equation}
These are genus-one sewing constraints.  Higher-genus sewing imposes further
compatibility between \(M\), chiral blocks, duality moves, and the pairing of
left and right chiral sectors.

\paragraph{Ising genus-one invariant.}
For the Ising chiral data of Section~\ref{sec:ising-chiral-theory},
restrict to full CFT
torus partition functions whose left and right chiral labels both lie in
\((\mathbf 1,\sigma,\varepsilon)\).  Let
\[
  M=(M_{ij})_{i,j\in\{\mathbf 1,\sigma,\varepsilon\}}
  \in \operatorname{Mat}_3(\mathbb Z_{\ge0})
\]
with \(M_{\mathbf 1\mathbf 1}=1\).  If \(M\) satisfies
\eqref{eq:modular-invariant-matrix} with
\eqref{eq:ising-modular-s} and \eqref{eq:ising-modular-t}, then \(M=I_3\).
Thus the only genus-one modular invariant with one vacuum in this chiral
Ising label set is the diagonal invariant
\begin{equation}
  Z_{\rm Ising}(\tau,\bar\tau)
  =
  |\chi_{\mathbf 1}|^2+|\chi_\sigma|^2+|\chi_\varepsilon|^2 .
  \label{eq:ising-diagonal-torus-partition-function}
\end{equation}
The \(T\)-constraint is already a spin-selection rule.  Since \(T\) is
diagonal,
\[
  (MT)_{ij}-(TM)_{ij}
  =
  M_{ij}\left(
    \ee^{2\pi\ii(h_j-c/24)}
    -
    \ee^{2\pi\ii(h_i-c/24)}
  \right).
\]
Therefore \(M_{ij}\ne0\) implies \(h_i-h_j\in\mathbb Z\).  For the Ising
weights
\[
  0,\qquad \frac1{16},\qquad \frac12,
\]
no two distinct weights differ by an integer.  Hence \(M\) is diagonal:
\[
  M=\operatorname{diag}(1,a,b),
  \qquad a,b\in\mathbb Z_{\ge0}.
\]
The \(S\)-constraint now gives, entry by entry,
\[
  (MS)_{\mathbf 1\sigma}=S_{\mathbf 1\sigma},
  \qquad
  (SM)_{\mathbf 1\sigma}=aS_{\mathbf 1\sigma}.
\]
Since \(S_{\mathbf 1\sigma}=1/\sqrt2\ne0\), one gets \(a=1\).  Similarly
\[
  (MS)_{\mathbf 1\varepsilon}=S_{\mathbf 1\varepsilon},
  \qquad
  (SM)_{\mathbf 1\varepsilon}=bS_{\mathbf 1\varepsilon},
\]
and \(S_{\mathbf 1\varepsilon}=1/2\ne0\), so \(b=1\).  Thus
\(M=I_3\), and \eqref{eq:ising-diagonal-torus-partition-function} follows
from \eqref{eq:rational-torus-partition-function}.

\subsection{Torus One-Point Blocks and Verlinde Defect Lines}
\label{subsec:torus-one-point-blocks-defect-lines}

The partition function is the genus-one zero-point function.  A complete
genus-one discussion also needs one-point blocks and topological defect
lines.  Both are often drawn pictorially; here we spell out the algebraic
content because the assumptions are different from those used in the scalar
partition function.

Let \(V\) be a rational chiral theory satisfying
Definition~\ref{def:sewing-compatible-rational-chiral-theory}, and let
\(v\in V\) be homogeneous of conformal weight \(h_v\).  The zero mode
\[
  o(v):=v_{h_v-1}
\]
has \(L_0\)-degree zero.  The chiral torus one-point block on the simple
module \(M_i\) is
\begin{equation}
  F_i(v;\tau)
  :=
  \operatorname{Tr}_{M_i}
  o(v)\,q^{L_0-c/24}.
  \label{eq:chiral-torus-one-point-block}
\end{equation}
Only the zero mode can contribute to a trace over \(L_0\)-eigenspaces:
a mode of nonzero \(L_0\)-degree maps each finite-dimensional eigenspace to a
different eigenspace and therefore has zero diagonal trace.  For
\(v=\mathbf 1\), \(o(v)=1\) and \(F_i(\mathbf 1;\tau)=\chi_i(\tau)\).

\begin{quotedtheorem}[Zhu modular covariance of torus one-point blocks]
\label{qthm:zhu-one-point-modular-covariance}
Assume \(V\) is rational, \(C_2\)-cofinite, of CFT type, and has ordinary
simple modules \(M_i\).  For each homogeneous \(v\in V\), the finite vector
\((F_i(v;\tau))_i\) spans a representation of \(SL_2(\mathbb Z)\).  If \(v\)
is Virasoro primary of weight \(h_v\), then for
\(\gamma=\begin{psmallmatrix}a&b\\ c&d\end{psmallmatrix}\)
\begin{equation}
  F_i\!\left(v;\frac{a\tau+b}{c\tau+d}\right)
  =
  (c\tau+d)^{h_v}
  \sum_j \rho_v(\gamma)_{ij}F_j(v;\tau).
  \label{eq:zhu-one-point-modular-covariance}
\end{equation}
For \(v=\mathbf 1\), \(\rho_v\) is the character modular representation
used above.  This is the one-point-function part of Zhu's genus-one
modular-covariance theorem \cite{Zhu1996}.
\end{quotedtheorem}

The factor \((c\tau+d)^{h_v}\) is the conformal weight factor for the
change of torus coordinate.  The theorem input is the finite-dimensionality
and holomorphic modular closure of the trace functions.

\paragraph{Mechanism behind torus one-point covariance.}
The one-point theorem is the genus-one version of transporting a conformal
block on a once-punctured torus.  The point of the zero mode \(o(v)\) is that
it preserves \(L_0\)-eigenspaces; hence the trace
\[
  F_i(v;\tau)=
  {\rm Tr}_{M_i}o(v)q^{L_0-c/24}
\]
is a genuine finite-energy trace after summing over each \(L_0\)-eigenspace.
In Zhu's square-bracket coordinate, the inserted state is transported by the
coordinate change at the puncture.  For a Virasoro primary \(v\), this
transport is scalar:
\[
  v\longmapsto (c\tau+d)^{h_v}v
\]
under the modular coordinate change
\[
  (z,\tau)\longmapsto
  \left(\frac{z}{c\tau+d},\frac{a\tau+b}{c\tau+d}\right).
\]
For descendants, the same coordinate change also produces lower-descendant
terms involving the Schwarzian and the square-bracket Virasoro modes; those
terms are the source of the quasimodular \(G_2\)-mixing in genus-one Ward
recursions.  The primary hypothesis in
Theorem~\ref{qthm:zhu-one-point-modular-covariance} is exactly the condition
that this coordinate transport has a single leading weight factor.

The finite-dimensional closure is the same \(C_2\)-cofinite trace-recursion
mechanism used for characters, but with \(v\) kept as a marked insertion.
Moving a square-bracket field around the torus expresses traces with
\([-2]\)-descendants in terms of traces with finitely many lower
square-bracket descendants and modular forms.  Rationality identifies the
simple ordinary modules as a finite set of sectors, while \(C_2\)-cofiniteness
keeps the marked-state quotient finite.  Therefore the traces
\[
  \{F_i(w;\tau):\ i\in I,\ w\in W_v\}
\]
for the finite descendant span \(W_v\) determined by \(v\) satisfy a finite
modular differential system.  Zhu's analytic theorem supplies convergence on
\(\mathbb H\), continuation of this finite solution space, and its invariance
under \(S\) and \(T\).  Formula
\eqref{eq:zhu-one-point-modular-covariance} is the primary, one-marked-state
component of that finite system.

A full-CFT torus one-point function of a bulk primary \(\Phi\) of weights
\((h,\bar h)\) is a sesquilinear pairing of the corresponding chiral
one-point-block vectors.  Modular covariance therefore requires an invariant
coefficient tensor for
\(\rho_\Phi(\gamma)\otimes \overline{\rho_\Phi(\gamma)}\), together with the
weight factor \((c\tau+d)^h(c\bar\tau+d)^{\bar h}\).  Thus torus one-point
functions constrain bulk OPE data beyond the scalar partition function.

Topological defect lines give a second genus-one probe.  In a diagonal
rational full CFT built from a unitary modular tensor category
\(\mathcal C\), the Verlinde defects are labelled by simple objects
\(a\in I\).  Let
\[
  \mathcal H_{\rm diag}
  =
  \bigoplus_{i\in I}M_i\otimes \overline{M_i^\vee}
\]
be the diagonal Hilbert space.  Define \(D_a\) on the summand labelled by
\(i\) by
\begin{equation}
  D_a\big|_{M_i\otimes \overline{M_i^\vee}}
  =
  \lambda_a(i)\operatorname{id},
  \qquad
  \lambda_a(i):=\frac{S_{ai}}{S_{0i}} .
  \label{eq:verlinde-defect-eigenvalue}
\end{equation}
The denominator is nonzero in a unitary modular tensor category because
\(S_{0i}=d_i S_{00}\) and \(d_i>0\).

\paragraph{Verlinde defect fusion and modular covariance.}
For the diagonal rational full CFT above,
\begin{equation}
  D_aD_b=\sum_c N_{ab}{}^cD_c .
  \label{eq:verlinde-defect-fusion}
\end{equation}
Moreover, the torus trace with \(D_a\) wrapping the Euclidean-time cycle is
\begin{equation}
  Z_a^{\rm time}(\tau,\bar\tau)
  =
  \sum_i \lambda_a(i)\chi_i(\tau)\overline{\chi_i(\tau)} .
  \label{eq:temporal-verlinde-defect-torus}
\end{equation}
After the modular \(S\)-move it becomes the spatial-defect trace
\begin{equation}
  Z_a^{\rm time}(-1/\tau,-1/\bar\tau)
  =
  \sum_{p,q}N_{ap}{}^q\,\chi_p(\tau)\overline{\chi_q(\tau)} .
  \label{eq:spatial-verlinde-defect-torus}
\end{equation}
The Verlinde formula says that the matrices \(N_a\) are simultaneously
diagonalized by \(S\), with eigenvalue \(S_{ai}/S_{0i}\) on the \(i\)-th
character.  Hence for every \(i\),
\[
  \lambda_a(i)\lambda_b(i)
  =
  \sum_c N_{ab}{}^c\lambda_c(i),
\]
which proves \eqref{eq:verlinde-defect-fusion} on each diagonal summand of
\(\mathcal H_{\rm diag}\).

For the modular transform, insert the character transformation law into
\eqref{eq:temporal-verlinde-defect-torus}.  The coefficient of
\(\chi_p\overline{\chi_q}\) is
\[
  \sum_i
  \frac{S_{ai}}{S_{0i}}\,S_{ip}\,\overline{S_{iq}}
  =
  \sum_i
  \frac{S_{ai}S_{pi}S_{q^\vee i}}{S_{0i}}
  =
  N_{ap}{}^q,
\]
where unitarity gives \(\overline{S_{iq}}=S_{q^\vee i}\), and the last
equality is Verlinde.  This gives
\eqref{eq:spatial-verlinde-defect-torus}.  The right side is a Hilbert-space
trace over
\[
  \mathcal H_a
  =
  \bigoplus_{p,q}N_{ap}{}^q\,
  M_p\otimes\overline{M_q^\vee},
\]
so the modular move has turned a temporal defect operator into a spatial
defect sector.

For the Ising labels \((\mathbf 1,\sigma,\varepsilon)\), the eigenvalue
vectors \((\lambda_a(\mathbf 1),\lambda_a(\sigma),\lambda_a(\varepsilon))\)
are
\[
  \lambda_{\mathbf 1}=(1,1,1),\qquad
  \lambda_\sigma=(\sqrt2,0,-\sqrt2),\qquad
  \lambda_\varepsilon=(1,-1,1).
\]
Thus
\[
  D_\varepsilon^2=D_{\mathbf 1},\qquad
  D_\varepsilon D_\sigma=D_\sigma,\qquad
  D_\sigma^2=D_{\mathbf 1}+D_\varepsilon,
\]
so \(D_\varepsilon\) is the invertible spin-flip line and \(D_\sigma\) is the
noninvertible Kramers--Wannier line.  Their modular covariance is completely
finite.  For example,
\[
  Z_\varepsilon^{\rm time}
  =
  |\chi_{\mathbf 1}|^2-|\chi_\sigma|^2+|\chi_\varepsilon|^2
\]
is sent by \(S\) to
\[
  \chi_{\mathbf 1}\overline{\chi_\varepsilon}
  +\chi_\varepsilon\overline{\chi_{\mathbf 1}}
  +|\chi_\sigma|^2,
\]
the Hilbert space twisted by the spin-flip defect.  Similarly
\[
  Z_\sigma^{\rm time}
  =
  \sqrt2\left(|\chi_{\mathbf 1}|^2-|\chi_\varepsilon|^2\right)
\]
is sent by \(S\) to
\[
  \chi_{\mathbf 1}\overline{\chi_\sigma}
  +\chi_\varepsilon\overline{\chi_\sigma}
  +\chi_\sigma\overline{\chi_{\mathbf 1}}
  +\chi_\sigma\overline{\chi_\varepsilon}.
\]
The same line-sliding argument gives selection rules for torus one-point
functions.  Since \(D_\varepsilon\) acts by \(-1\) on the Ising spin primary,
the torus one-point function of \(\sigma\) equals its negative and therefore
vanishes.  This topological-defect Ward identity is independent of any
particular lattice realization.

For a unitary compact two-dimensional CFT with discrete spectrum on the
circle, the rectangular torus partition function
\[
  Z(\beta)=\operatorname{Tr}_{\mathcal H(S^1)}
  \exp\{-\beta(H-c_{\rm tot}/24)\},
  \qquad c_{\rm tot}=c+\bar c,
\]
obeys modular invariance
\[
  Z(\beta)=Z(4\pi^2/\beta)
\]
when the spatial circle has circumference \(2\pi\).  If the low-temperature
trace is dominated by a unique vacuum, then as \(\beta\to0^+\),
\begin{equation}
  \log Z(\beta)
  =
  \frac{\pi^2 c_{\rm tot}}{6\beta}+o(\beta^{-1}).
  \label{eq:cardy-partition-leading}
\end{equation}
The derivation is only the modular transformation:
low temperature at \(4\pi^2/\beta\) gives
\[
  Z(4\pi^2/\beta)
  =
  \exp\left\{\frac{4\pi^2}{\beta}\frac{c_{\rm tot}}{24}\right\}
  (1+o(1)),
\]
which is \eqref{eq:cardy-partition-leading}.  Tauberian hypotheses convert
this high-temperature partition-function statement into an asymptotic density
of states.  Those hypotheses are analytic input; modular invariance alone is
not a pointwise formula for individual degeneracies.

\begin{quotedtheorem}[Exponential Tauberian theorem]
\label{qthm:exponential-tauberian-cardy}
Let \(N(E)\) be a nondecreasing right-continuous function on
\([0,\infty)\), and suppose its Laplace transform
\[
  Z(\beta)=\int_{[0,\infty)}\ee^{-\beta E}\,\dd N(E)
\]
is finite for every \(\beta>0\).  If, for some \(A>0\),
\[
  \log Z(\beta)=\frac A\beta+o(\beta^{-1})
  \qquad(\beta\downarrow0),
\]
then
\[
  \log N(E)=2\sqrt{AE}+o(\sqrt E)
  \qquad(E\to\infty).
\]
This is the exponential Tauberian input used in Cardy asymptotics; the
regular-variation reference is a source trace, while the chapter below spells
out the one-sided Laplace estimate and isolates the inverse Tauberian step
\cite{BinghamGoldieTeugels1987}.
\end{quotedtheorem}

The theorem contains an elementary one-sided estimate and a genuinely
Tauberian inverse estimate.  Let \(\mu_N\) be the positive Stieltjes measure
with \(\mu_N([0,E])=N(E)\).  For every \(\beta>0\),
\[
  Z(\beta)
  \ge
  \int_{[0,E]}\ee^{-\beta E}\,\dd\mu_N(E')
  =
  \ee^{-\beta E}N(E),
\]
and hence
\begin{equation}
  \log N(E)
  \le
  \inf_{\beta>0}\{\beta E+\log Z(\beta)\}.
  \label{eq:cardy-tauberian-laplace-upper-bound}
\end{equation}
If \(\log Z(\beta)=A/\beta+o(\beta^{-1})\), choosing
\(\beta=\sqrt{A/E}\) in \eqref{eq:cardy-tauberian-laplace-upper-bound}
gives the upper half
\[
  \limsup_{E\to\infty}{\frac{\log N(E)}{\sqrt E}}\le 2\sqrt A .
\]
The inverse half is the nontrivial content: positivity and monotonicity of the
counting measure force enough spectral mass in the saddle window
\(E'\sim A/\beta^2\) to produce the same leading exponent from below.  Without
that positivity, cancellations in a signed trace can obey the same modular
transformation law while carrying no comparable counting interpretation.

The coefficient \(2\sqrt A\) is fixed by the Legendre saddle.  If a candidate
cumulative growth has the leading form \(\log N(E)\sim B\sqrt E\), the formal
Laplace exponent is
\[
  \Phi_\beta(E')=B\sqrt{E'}-\beta E',
  \qquad
  E'_\beta=\frac{B^2}{4\beta^2},
\]
and
\[
  \sup_{E'\ge0}\Phi_\beta(E')=\frac{B^2}{4\beta}.
\]
Matching this with \(A/\beta\) gives \(B=2\sqrt A\).  The quoted theorem is
precisely the analytic statement that this saddle reasoning is legitimate for
a positive cumulative spectral measure at the exponential scale.

This theorem is the precise analytic bridge between the modular
high-temperature statement and the density of states.  Its positivity
hypothesis is essential: \(N(E)\) is a cumulative spectral counting function,
not a signed distribution obtained by subtracting sectors.

\paragraph{CFT-internal Cardy growth.}
Let \(\mathcal C\) be a unitary compact two-dimensional CFT on a circle of
circumference \(2\pi\).  Assume the Hilbert-space spectrum of
\[
  E=L_0+\bar L_0
\]
is discrete with finite multiplicities, \(E\ge0\), and has a unique vacuum at
\(E=0\).  Assume also that the torus partition function is modular invariant
and has no continuous spectral contribution.  Let
\[
  N(E)=\dim\{\psi\in\mathcal H(S^1)\mid
  (L_0+\bar L_0)\psi\le E\}
\]
be the cumulative degeneracy.  Then
\begin{equation}
  \log N(E)
  =
  2\pi\sqrt{\frac{c_{\rm tot}E}{6}}+o(\sqrt E),
  \qquad
  E\to\infty .
  \label{eq:cardy-cumulative-density}
\end{equation}
This is a direct consequence of the quoted Tauberian theorem after identifying
the positive counting measure.  The shifted cylinder partition function used
above is
\[
  Z(\beta)
  =
  \operatorname{Tr}\exp\{-\beta(L_0+\bar L_0-c_{\rm tot}/24)\}.
\]
Multiplying by \(\exp\{-\beta c_{\rm tot}/24\}\) gives the ordinary Laplace
transform of the positive counting measure \(\dd N(E)\):
\[
  \widetilde Z(\beta)
  =
  \operatorname{Tr}\exp\{-\beta(L_0+\bar L_0)\}
  =
  \int_{[0,\infty)}\ee^{-\beta E}\,\dd N(E).
\]
Since \(\beta c_{\rm tot}/24=o(\beta^{-1})\), the high-temperature asymptotic
\eqref{eq:cardy-partition-leading} gives
\[
  \log \widetilde Z(\beta)
  =
  \frac{\pi^2c_{\rm tot}}{6\beta}+o(\beta^{-1}).
\]
Apply Theorem~\ref{qthm:exponential-tauberian-cardy} with
\(A=\pi^2c_{\rm tot}/6\).

A saddle calculation checks the coefficient but does not replace the
Tauberian theorem or the positivity hypothesis above.  With
\(A=\pi^2c_{\rm tot}/6\), if a smooth density had entropy
\[
  S(E)=2\sqrt{AE},
\]
then the Laplace exponent \(S(E)-\beta E\) is stationary at
\[
  E_\ast=\frac A{\beta^2},
\]
and the stationary value is \(A/\beta\).  Thus the density-growth coefficient
in \eqref{eq:cardy-cumulative-density} is exactly the Legendre dual of the
modular coefficient in \eqref{eq:cardy-partition-leading}; the theorem above
is what justifies this conclusion for a discrete positive spectrum.

The modular bootstrap is the study of the constraints
\eqref{eq:modular-invariant-matrix}, reflection positivity, spin integrality,
and higher-genus sewing on the spectrum and OPE data.  In rational theories it
is finite-dimensional algebra.  In nonrational compact theories it is an
infinite-dimensional positivity and growth problem for torus and higher-genus
partition functions.

\section{Direct-Integral Data for Nonrational CFT}
\label{sec:nonrational-cft-direct-integral-data}

The rational theory discussed above has two simplifying properties: the
chiral sector category has finitely many simple objects, and every sewing
formula reduces to a finite sum.  Neither property is part of the local
definition of a two-dimensional conformal field theory.  A compact CFT may
have a countably infinite discrete spectrum, as in an irrational compact
boson.  A noncompact CFT may have continuous normalizable spectrum, as in
Liouville theory.  In both cases the finite sum over intermediate primaries
must be replaced by a spectral resolution of Hilbert-space vectors.

\begin{definition}[Spectral radial product datum]
\label{def:spectral-radial-product-datum}
Fix a unitary two-dimensional CFT on the round circle with Hilbert space
\(\mathcal H\), vacuum \(\Omega\), self-adjoint nonnegative operators
\(L_0,\bar L_0\), and dense local domain \(\mathcal D_{\rm loc}\).  A
spectral radial product datum for two local fields \(A,B\) consists of the
following data.
\begin{enumerate}
  \item For every \(0<r<1\) and \(\theta\in\mathbb R/2\pi\mathbb Z\), the
  radially ordered product vector
  \[
    \Phi_{BA}(r,\theta)=B(re^{\ii\theta})A(0)\Omega
  \]
  is a vector in \(\mathcal H\).
  \item Let \(\mathcal H_{BA}\subset\mathcal H\) be the closed subspace
  generated by all \(\Phi_{BA}(r,\theta)\) and by the action of the two
  Virasoro algebras.  The representation of the commuting spectral observables
  used to label intermediate sectors is represented as a direct integral
  \begin{equation}
    \mathcal H_{BA}
    \simeq
    \int_{X_{BA}}^{\oplus}\mathcal K_x\,\dd\mu_{BA}(x),
    \label{eq:direct-integral-ba}
  \end{equation}
  where \(X_{BA}\) is a standard Borel space, \(\mu_{BA}\) is a positive
  measure, and \(x\mapsto\mathcal K_x\) is a measurable field of Hilbert
  spaces carrying the descendant and multiplicity degrees of freedom.
  \item In this representation the product vector is a square-integrable
  measurable vector field
  \[
    \widehat\Phi_{BA}(x;r,\theta)\in\mathcal K_x,
    \qquad
    \int_{X_{BA}}\|\widehat\Phi_{BA}(x;r,\theta)\|_{\mathcal K_x}^2
      \dd\mu_{BA}(x)<\infty .
  \]
\end{enumerate}
The variable \(x\) is an intermediate spectral label, not necessarily the
label of a normalizable primary vector.  When \(x\) lies in a continuous
spectrum, the formal object called a primary of label \(x\) is a generalized
eigenvector; the Hilbert-space object is the square-integrable field
\(\widehat\Phi_{BA}\).
\end{definition}

This definition is deliberately phrased before choosing a basis of chiral
blocks.  The direct integral is supplied by the spectral theorem for the
commuting self-adjoint operators and by the decomposition of the chiral
representation on the closed cyclic subspace.  A conformal-block expansion is
obtained only after choosing measurable bases of intertwiners in the fibers
\(\mathcal K_x\).  The Plancherel measure and the multiplicity fields are
therefore part of the CFT data.

\begin{remark}[Direct-integral OPE identity]
\label{rem:direct-integral-ope-identity}
Let \(A,B,C,D\in\mathcal D_{\rm loc}\) and assume that the radial products
\(\Phi_{BA}(r,\theta)\) and \(\Phi_{CD}(\rho,\varphi)\) have spectral radial
product data using a common direct-integral decomposition
\[
  \mathcal H_{ABCD}
  \simeq
  \int_X^{\oplus}\mathcal K_x\,\dd\mu(x)
\]
of the closed subspace generated by the two product families.  Then the
four-point function in the radial domain \(0<r<\rho<1\) is
\begin{equation}
  \langle
    C(\rho e^{\ii\varphi})D(0)\Omega,\,
    B(re^{\ii\theta})A(0)\Omega
  \rangle_{\mathcal H}
  =
  \int_X
  \left\langle
    \widehat\Phi_{CD}(x;\rho,\varphi),\,
    \widehat\Phi_{BA}(x;r,\theta)
  \right\rangle_{\mathcal K_x}
  \dd\mu(x).
  \label{eq:direct-integral-four-point}
\end{equation}
If the measure is counting measure on a finite set this is the rational
finite sum over intermediate states.  If the measure has a continuous part,
the same formula is an \(L^2\)-inner product, not a pointwise sum over
delta-normalized states.
Indeed, choose the unitary map
\[
  U:\mathcal H_{ABCD}\longrightarrow
  \int_X^{\oplus}\mathcal K_x\,\dd\mu(x)
\]
implementing the direct integral.  By definition,
\(U\Phi_{BA}=\widehat\Phi_{BA}\) and
\(U\Phi_{CD}=\widehat\Phi_{CD}\).  Unitarity of \(U\) gives the Plancherel
identity
\[
  \langle \Phi_{CD},\Phi_{BA}\rangle_{\mathcal H}
  =
  \int_X
    \langle
      (U\Phi_{CD})(x),(U\Phi_{BA})(x)
    \rangle_{\mathcal K_x}\dd\mu(x),
\]
which is \eqref{eq:direct-integral-four-point}.  No assumption about a
discrete basis of normalizable primaries has entered the argument.
\end{remark}

The usual OPE notation is a compressed way of writing
\eqref{eq:direct-integral-four-point}.  For a continuous spectrum it should
be read as an identity of matrix elements after smearing the intermediate
label against the Plancherel measure and after specifying the domain of
radial convergence.

\subsection{Worked Example: The Noncompact Free Boson}
\label{subsec:noncompact-free-boson-direct-integral}

The simplest continuous-spectrum example is a single real free boson with
target \(\mathbb R\).  We use the same chiral normalization as in
Section~\ref{sec:toroidal-narain-lattice-cfts},
\[
  X_L(z)X_L(0)\sim-\log z,\qquad
  X_R(\bar z)X_R(0)\sim-\log\bar z.
\]
The zero-mode Hilbert space is \(L^2(\mathbb R,\dd x)\).  After Fourier
transform it is \(L^2(\mathbb R,\dd p)\), and the full Hilbert space is the
direct integral
\begin{equation}
  \mathcal H_{\mathbb R}
  =
  \int_{\mathbb R}^{\oplus}
  \mathcal F_{p}\,\dd p,
  \qquad
  \mathcal F_p\simeq\mathcal F_{\rm osc}^L\otimes
  \mathcal F_{\rm osc}^R .
  \label{eq:noncompact-boson-hilbert-direct-integral}
\end{equation}
The fiber ground state at momentum \(p\) has
\begin{equation}
  h_p=\bar h_p=\frac{p^2}{2}.
  \label{eq:noncompact-boson-weights}
\end{equation}
Thus \(p\) is a spectral label in the direct integral.  The plane wave
\(\ee^{\ii p x}\) is a generalized eigenvector; an actual Hilbert vector is a
square-integrable wave packet \(f(p)\in L^2(\mathbb R,\dd p)\), together with
oscillator excitations.

For \(a\in\mathbb R\), let \(V_a(z,\bar z)\) denote the diagonal exponential
field
\[
  V_a(z,\bar z)
  =
  :\exp\!\left(\ii aX_L(z)+\ii aX_R(\bar z)\right): .
\]
As a zero-mode operator it translates the momentum fiber,
\[
  V_a:\mathcal F_p\longrightarrow \mathcal F_{p+a}.
\]
The field at a fixed real label \(a\) is distributional in the spectral
label.  For \(f\in C_c^\infty(\mathbb R)\), the smeared exponential
\begin{equation}
  V_f(z,\bar z)
  =
  \int_{\mathbb R} f(a)V_a(z,\bar z)\,\dd a
  \label{eq:noncompact-boson-smeared-vertex}
\end{equation}
has finite matrix elements between finite-energy wave packets in every
closed annulus avoiding coincident points.

The leading product of two labelled exponentials is
\begin{equation}
  V_a(z,\bar z)V_b(0)
  \sim
  |z|^{2ab}V_{a+b}(0)+\text{oscillator descendants}.
  \label{eq:noncompact-boson-labelled-ope}
\end{equation}
Consequently the smeared leading OPE coefficient is the weighted convolution
\begin{equation}
  V_f(z,\bar z)V_g(0)
  \sim
  \int_{\mathbb R}
  (f\star_z g)(c)\,V_c(0)\,\dd c+\cdots ,
  \label{eq:noncompact-boson-smeared-ope}
\end{equation}
where
\begin{equation}
  (f\star_z g)(c)
  =
  \int_{\mathbb R}
  f(a)g(c-a)|z|^{2a(c-a)}\,\dd a .
  \label{eq:noncompact-boson-weighted-convolution}
\end{equation}
This is the continuous analogue of the lattice charge-conservation rule in
\eqref{eq:narain-vertex-ope-leading}.  Associativity of the leading charge
factor is the elementary identity
\begin{equation}
  ab+(a+b)c=bc+a(b+c),
  \label{eq:noncompact-boson-ope-associativity-exponent}
\end{equation}
which says that either radial grouping of \(V_aV_bV_c\) gives the same
charge \(a+b+c\) and the same power of pairwise distances after the standard
coordinate factors are restored.  The full diagonal field is separated-point
single-valued because the left and right monodromy phases cancel.

The genus-one zero-point function is a trace density.  If one puts the target
coordinate in a box of length \(L\), the ordinary trace is proportional to
\(L\).  Dividing by \(L\) and letting \(L\to\infty\) gives the density
\begin{equation}
  Z_{\mathbb R}^{\rm dens}(\tau,\bar\tau)
  =
  \frac{1}{|\eta(\tau)|^2}
  \int_{\mathbb R}
  \exp(-2\pi\tau_2p^2)\,\dd p
  =
  \frac{1}{\sqrt{2\tau_2}\,|\eta(\tau)|^2},
  \qquad
  \tau_2=\operatorname{Im}\tau .
  \label{eq:noncompact-boson-partition-density}
\end{equation}
The equality uses the ordinary Gaussian integral in the chosen Plancherel
normalization.  The density is modular invariant: under
\(S:\tau\mapsto-1/\tau\), \(\tau_2\mapsto\tau_2/|\tau|^2\) while
\(|\eta(-1/\tau)|^2=|\tau|\,|\eta(\tau)|^2\), so the factor
\(\tau_2^{-1/2}|\eta(\tau)|^{-2}\) is unchanged.  This is a statement about
the semifinite trace per unit target volume; it is the correct replacement
for a finite Hilbert-space trace in a translation-invariant noncompact
target.

The same density is obtained from the compact circle by decompactification.
For a circle of radius \(R\), the momentum spacing is \(1/R\).  At fixed
\(\tau_2>0\), the winding sectors with nonzero winding have weights
exponentially small as \(R\to\infty\), and
\[
  \frac1R\sum_{m\in\mathbb Z}
  \exp\!\left(-2\pi\tau_2\frac{m^2}{R^2}\right)
  \longrightarrow
  \int_{\mathbb R}\exp(-2\pi\tau_2p^2)\,\dd p .
\]
Thus compact-boson theta functions converge, after division by the target
volume normalization, to the Plancherel density
\eqref{eq:noncompact-boson-partition-density}.

This example isolates the continuous-spectrum mechanics from the additional
analytic content of Liouville theory.  In the free noncompact boson the
Plancherel measure is Lebesgue measure, the reflection relation is absent,
and the fusing transform is the Fourier/Plancherel transform attached to
linear momentum conservation.  In Liouville theory the same direct-integral
language remains, but the spectrum is
\(\alpha=Q/2+\ii P\) modulo reflection, the three-point coefficient is DOZZ,
and the fusing transform is the Virasoro noncompact fusion kernel.  The
difference is data, not notation.

\begin{definition}[Measurable fusing transform]
\label{def:measurable-fusing-transform}
For fixed ordered external fields \(A,B,C,D\), let
\[
  \int_{X_s}^{\oplus}\mathcal K_x^{(s)}\,\dd\mu_s(x),
  \qquad
  \int_{X_t}^{\oplus}\mathcal K_y^{(t)}\,\dd\mu_t(y)
\]
be the direct-integral Hilbert spaces obtained from the \(s\)-channel
radial grouping \((BA)(DC)\) and the \(t\)-channel grouping \((CB)(DA)\).
A measurable fusing transform is a unitary operator
\begin{equation}
  \mathbb F_{st}:
  \int_{X_s}^{\oplus}\mathcal K_x^{(s)}\,\dd\mu_s(x)
  \longrightarrow
  \int_{X_t}^{\oplus}\mathcal K_y^{(t)}\,\dd\mu_t(y)
  \label{eq:measurable-fusing-transform}
\end{equation}
which intertwines the two analytic continuations of the same chiral
four-point block from the \(s\)- and \(t\)-channel radial domains.  When the
fibers are finite-dimensional and a measurable basis has been chosen, this
operator is represented by a kernel
\begin{equation}
  (\mathbb F_{st} f)(y)
  =
  \int_{X_s}
  F_{st}(y,x)f(x)\,\dd\mu_s(x),
  \label{eq:fusing-kernel-direct-integral}
\end{equation}
understood as an equality in the direct-integral Hilbert space, with the
unitarity condition
\begin{equation}
  \int_{X_t} F_{st}(y,x)^\dagger F_{st}(y,x')\,\dd\mu_t(y)
  =
  \delta_{\mu_s}(x,x')\,\mathbf 1
  \label{eq:fusing-kernel-unitarity}
\end{equation}
in distributional notation.
\end{definition}

\paragraph{Plancherel form of crossing.}
Assume the hypotheses of Definition~\ref{def:measurable-fusing-transform}.
Let \(v_s,w_s\) be the two \(s\)-channel square-integrable coefficient fields
whose inner product gives the four-point function in an \(s\)-channel radial
domain, and let \(v_t,w_t\) be the corresponding \(t\)-channel coefficient
fields.  If
\begin{equation}
  v_t=\mathbb F_{st}v_s,\qquad w_t=\mathbb F_{st}w_s,
  \label{eq:coefficient-fields-fusing-transform}
\end{equation}
then unitarity of the fusing transform gives
\begin{equation}
  \int_{X_s}
    \langle w_s(x),v_s(x)\rangle_{\mathcal K_x^{(s)}}\,\dd\mu_s(x)
  =
  \int_{X_t}
    \langle w_t(y),v_t(y)\rangle_{\mathcal K_y^{(t)}}\,\dd\mu_t(y).
  \label{eq:direct-integral-crossing}
\end{equation}
In this formulation the displayed Hilbert-space identity is the final
Plancherel step.  The mathematical content is the construction of the
direct-integral decompositions, the proof that the conformal blocks admit the
unitary transform \(\mathbb F_{st}\), and the verification of
\eqref{eq:coefficient-fields-fusing-transform} for the actual OPE kernels.
In a rational CFT the same statement reduces to a finite matrix identity.  In
Liouville theory the label spaces are continuous and the identity becomes the
Ponsot--Teschner Virasoro fusion-kernel identity, together with the DOZZ
coefficient relation.

\begin{definition}[Nonrational genus-zero bootstrap problem]
\label{def:nonrational-genus-zero-bootstrap-problem}
A nonrational genus-zero bootstrap problem consists of:
\begin{enumerate}
  \item a class of local fields and their dense radial-product domain;
  \item for every ordered pair of fields, direct-integral intermediate
  spectra with Plancherel measures and multiplicity fibers;
  \item measurable three-point/OPE coefficient fields;
  \item chiral conformal blocks with specified radial convergence domains
  and analytic continuations;
  \item measurable fusing and braiding transforms satisfying unitarity,
  pentagon, and hexagon identities as identities of direct-integral
  operators;
  \item a reflection-positive pairing of left and right chiral data producing
  single-valued Euclidean correlation distributions.
\end{enumerate}
Solving this problem means proving existence, uniqueness up to unitary
equivalence of the resulting correlation theory, reflection positivity,
locality, and compatibility with sewing in the domains under consideration.
Crossing equations for individual four-point functions are necessary
consequences of such a solution, not a replacement for the direct-integral
and sewing data.
\end{definition}

\begin{example}[Two spectral types]
\label{ex:nonrational-cft-two-spectral-types}
For an irrational compact free boson the intermediate label space is
countable, the measure is counting measure, and modular covariance is governed
by theta functions and Poisson summation rather than by a finite modular
matrix.  For Liouville theory the normalizable bulk spectrum is
\[
  \alpha=\frac Q2+\ii P,\qquad P\in\mathbb R_{\ge0},
\]
with Plancherel measure fixed by the two-point normalization and reflection
amplitude.  The four-point function is the direct integral displayed in
Definition~\ref{def:liouville-continuous-blocks}.  These two examples have
different spectral measures, but both are governed by the same Hilbert-space
identity \eqref{eq:direct-integral-four-point}.
\end{example}

\section{Logarithmic Theories}

A logarithmic CFT is a conformal field theory whose chiral module category is
not semisimple and in which the energy operator \(L_0\) need not be
diagonalizable.  Diagonal \(L_0\) is what made the rational characters honest
sums of powers of \(q\)
over simple modules; a nilpotent part of \(L_0\) changes both local conformal
covariance and the genus-one trace problem.

\begin{definition}[Rank-two logarithmic Virasoro pair]
\label{def:rank-two-logarithmic-pair}
A rank-two logarithmic pair of chiral fields of weight \(h\) is a pair
\((\mathcal C,\mathcal D)\) whose states \(C,D\) satisfy
\[
  L_0C=hC,\qquad L_0D=hD+C,\qquad L_nC=L_nD=0\quad(n>0).
\]
Equivalently, the singular part of the stress-tensor OPE is
\[
  T(\zeta)\mathcal C(z)
  \sim
  \frac{h\mathcal C(z)}{(\zeta-z)^2}
  +
  \frac{\partial\mathcal C(z)}{\zeta-z},
\]
\[
  T(\zeta)\mathcal D(z)
  \sim
  \frac{h\mathcal D(z)+\mathcal C(z)}{(\zeta-z)^2}
  +
  \frac{\partial\mathcal D(z)}{\zeta-z}.
\]
Thus \(\mathcal C\) is primary, while \(\mathcal D\) is its logarithmic
partner.  With \(N D=C\) and \(NC=0\), the two-dimensional span of
\((C,D)\) carries \(L_0=h\operatorname{id}+N\).
\end{definition}

The finite consequences are immediate and worth recording because this is
where signs and \(2\pi\ii\)'s enter.  On states,
\[
  q^{L_0-c/24}C=q^{h-c/24}C,\qquad
  q^{L_0-c/24}D=q^{h-c/24}\bigl(D+2\pi\ii\tau\,C\bigr),
  \qquad q=\ee^{2\pi\ii\tau}.
\]
On fields, the Virasoro commutators are, for \(n=-1,0,1\),
\[
  [L_n,\mathcal C(z)]
  =
  \bigl(z^{n+1}\partial_z+(n+1)hz^n\bigr)\mathcal C(z),
\]
\[
  [L_n,\mathcal D(z)]
  =
  \bigl(z^{n+1}\partial_z+(n+1)hz^n\bigr)\mathcal D(z)
  +(n+1)z^n\mathcal C(z).
\]
In particular,
\[
  \ee^{sL_0}\mathcal C(z)\ee^{-sL_0}
  =
  \ee^{sh}\mathcal C(\ee^s z),
\]
\[
  \ee^{sL_0}\mathcal D(z)\ee^{-sL_0}
  =
  \ee^{sh}\bigl(\mathcal D(\ee^s z)+s\mathcal C(\ee^s z)\bigr).
\]
The additive \(s=\log\lambda\) term is the local origin of logarithms in
correlators.

\begin{proposition}[Two-point functions of a logarithmic pair]
\label{prop:logarithmic-pair-two-point-functions}
Assume the vacuum is invariant under \(L_{-1},L_0,L_1\), the two fields in
Definition~\ref{def:rank-two-logarithmic-pair} are mutually local, and their
two-point functions are single-valued on a chosen simply connected
punctured coordinate chart after a branch of \(\log z\) is fixed.  Then
translation, scaling, and special-conformal Ward identities force
\[
  \langle\mathcal C(z)\mathcal C(0)\rangle=0,
\]
\[
  \langle\mathcal C(z)\mathcal D(0)\rangle
  =
  b\,z^{-2h},
\]
\[
  \langle\mathcal D(z)\mathcal D(0)\rangle
  =
  (d-2b\log z)z^{-2h},
\]
for constants \(b,d\in\mathbb C\).  The constant \(d\) depends on the choice
of logarithmic partner \(D\), while \(b\) is invariant under
\(D\mapsto D+\lambda C\).
\end{proposition}

\begin{proof}
Translation invariance makes the correlators functions of
\(z=z_1-z_2\).  Write
\[
  G_{CC}(z)=\langle\mathcal C(z)\mathcal C(0)\rangle,\qquad
  G_{CD}(z)=\langle\mathcal C(z)\mathcal D(0)\rangle,
\]
\[
  G_{DD}(z)=\langle\mathcal D(z)\mathcal D(0)\rangle.
\]
The \(L_0\) Ward identity gives
\[
  (z\partial_z+2h)G_{CC}=0,
\]
\[
  (z\partial_z+2h)G_{CD}+G_{CC}=0,
\]
\[
  (z\partial_z+2h)G_{DD}+2G_{CD}=0.
\]
Hence, before imposing \(L_1\),
\[
  G_{CC}=A z^{-2h},\qquad
  G_{CD}=(B-A\log z)z^{-2h},
\]
\[
  G_{DD}=(D_0-2B\log z+A(\log z)^2)z^{-2h}.
\]
This calculation is the scale-covariant source of logarithms.

Now impose the \(L_1\) Ward identity on
\(\langle\mathcal C(z_1)\mathcal D(z_2)\rangle\).  The ordinary primary part
of the differential operator is
\[
  (z_1+z_2)(z\partial_z+2h)G_{CD}(z),
  \qquad z=z_1-z_2,
\]
and the nilpotent term from the \(\mathcal D\)-insertion is
\[
  2z_2G_{CC}(z).
\]
Using the \(L_0\) equation, the total \(L_1\) variation is
\[
  -(z_1+z_2)G_{CC}(z)+2z_2G_{CC}(z)
  =
  -z\,G_{CC}(z).
\]
It must vanish for \(z\ne0\), so \(G_{CC}=0\) and therefore \(A=0\).  The
remaining \(L_0\) equations give the displayed forms with \(b=B\) and
\(d=D_0\).  The \(L_1\) equation for \(G_{DD}\) is then automatic:
the primary differential part gives
\(-2b(z_1+z_2)z^{-2h}\), while the two nilpotent insertions contribute
\(2b(z_1+z_2)z^{-2h}\).

Finally replace \(D\) by \(D+\lambda C\).  Since \(G_{CC}=0\),
\[
  G_{CD}\mapsto G_{CD},\qquad
  G_{DD}\mapsto G_{DD}+2\lambda G_{CD}.
\]
Thus \(b\) is unchanged and \(d\) is shifted to \(d+2\lambda b\).
\end{proof}

The ordinary character sees only the trace of
\(q^{L_0-c/24}\).  On the span of \(C,D\),
\[
  \operatorname{Tr}\bigl(q^{L_0-c/24}\bigr)
  =
  2q^{h-c/24},
\]
because the nilpotent coefficient \(2\pi\ii\tau N\) has trace zero.  Thus an
ordinary character records the generalized eigenvalue and multiplicity but
not the extension class producing the logarithmic partner.  In a finite-length
\(C_2\)-cofinite logarithmic theory, the genus-one space closed under
modular transformations is therefore larger than the span of ordinary
characters: it includes generalized characters and pseudo-trace functions on
projective modules.  A pseudo-trace is the replacement for the usual matrix
trace when the endomorphism algebra of a projective module is not semisimple;
it is a linear functional with the cyclicity property needed for torus
sewing, and it can couple to the nilpotent part invisible to the ordinary
trace.

\paragraph{Finite projective pseudo-trace model.}
The finite algebraic mechanism is already visible without invoking the full
logarithmic VOA modular theorem.  Let \(A\) be a finite-dimensional
associative algebra over \(\mathbb C\), let \(P\) be a finitely generated
projective left \(A\)-module, and let
\[
  \varphi:A\longrightarrow\mathbb C
\]
be a symmetric linear functional, meaning
\[
  \varphi(ab)=\varphi(ba),\qquad a,b\in A .
\]
Choose a projective dual basis
\[
  p_r\in P,\qquad f^r\in\operatorname{Hom}_A(P,A),\qquad
  x=\sum_r p_r f^r(x)\quad (x\in P).
\]
For \(T\in\operatorname{End}_A(P)\), define the
\(\varphi\)-pseudo-trace
\begin{equation}
  \operatorname{ptr}_{P}^{\varphi}(T)
  :=
  \sum_r
  \varphi\bigl(f^r(Tp_r)\bigr).
  \label{eq:finite-projective-pseudo-trace}
\end{equation}
The expression is independent of the chosen dual basis.  This is the
Hattori--Stallings trace paired with \(\varphi\): changing dual bases changes
\(\sum_r f^r(Tp_r)\) by a commutator in \(A\), and the symmetric property of
\(\varphi\) kills commutators.  It follows in particular that
\[
  \operatorname{ptr}_{P}^{\varphi}(ST)
  =
  \operatorname{ptr}_{P}^{\varphi}(TS),
  \qquad
  S,T\in\operatorname{End}_A(P),
\]
which is the finite cyclicity property needed when the torus sewing trace is
rotated around a nonsemisimple channel.

The smallest example is the dual-number algebra
\[
  A=\mathbb C[\epsilon]/(\epsilon^2),
  \qquad
  P=A .
\]
Let \(\varphi_{\rm res}(a+b\epsilon)=b\).  Multiplication by
\(a+b\epsilon\) on the vector space with basis \((1,\epsilon)\) has ordinary
trace \(2a\), so the ordinary trace kills the nilpotent coefficient \(b\).
The projective pseudo-trace with \(p_1=1\) and \(f^1(x)=x\), however, gives
\[
  \operatorname{ptr}^{\varphi_{\rm res}}_{A}(R_{a+b\epsilon})
  =
  \varphi_{\rm res}(a+b\epsilon)
  =
  b ,
\]
where \(R_{a+b\epsilon}\) denotes right multiplication, an \(A\)-module
endomorphism of the regular left module.  If the Jordan cell of \(L_0\) is
modeled by \(L_0=h+\epsilon\), then
\[
  q^{L_0-c/24}
  =
  q^{h-c/24}\bigl(1+2\pi\ii\tau\,\epsilon\bigr),
\]
and the ordinary trace gives \(2q^{h-c/24}\), whereas the pseudo-trace gives
\[
  \operatorname{ptr}^{\varphi_{\rm res}}_{A}
  \bigl(q^{L_0-c/24}\bigr)
  =
  2\pi\ii\tau\,q^{h-c/24}.
\]
Thus the polynomial \(\tau\)-dependence that ordinary characters miss is not a
decorative correction; it is the projective trace of the nilpotent extension
data.  In logarithmic VOA applications the algebra \(A\) is replaced by
endomorphism algebras of projective covers or by the appropriate finite
quotients controlling torus one-point functions, and
\eqref{eq:finite-projective-pseudo-trace} is the finite model for the
pseudo-trace functions that enter modular closure.

The semisimple Verlinde formula
\eqref{eq:verlinde-formula} is consequently replaced by categorical
statements about the Grothendieck ring, projective modules, modified traces,
and the factorizable braided finite tensor category attached to the
logarithmic chiral theory.  The right question is not whether the same matrix
formula survives unchanged, but which categorical pairing diagonalizes which
quotient of the tensor product.  This is one of the places where the
algebraic formulation is essential: it names the objects whose semisimplicity
has failed and identifies the extra data needed to recover modular sewing.

\section{Output}

The output of this chapter is a hierarchy of definitions and hypotheses:
The construction has the following order.  A VOA defines modules and chiral
blocks.  Chiral blocks sew over the moduli space of pointed curves.  In the
semisimple finite case, sewing data assemble into a modular tensor category.
The full CFT is then obtained by choosing a compatible pairing of left and
right chiral theories, equivalently by imposing the modular-invariance and
factorization constraints on torus and higher-genus amplitudes.
The Ising model gives the finite rational example.  Logarithmic theories show
which parts of the rational construction depend on semisimplicity and which
parts persist after replacing simple-module linear algebra by finite tensor
category data.
