\chapter{Liouville Conformal Field Theory}
\label{ch:liouville-cft}
\ConventionsEuclideanCFT

\section{Classical Datum and Stress Tensor}

Liouville theory is the noncompact two-dimensional conformal theory whose
interaction is an exponential of a scalar field.  This chapter treats it as a
two-dimensional QFT construction problem whose data are a background surface,
a scalar distribution, a curvature coupling, exponential insertions, Virasoro
Ward identities, degenerate fields, and sewing constraints.  The development
below fixes the normalizations, derives the local representation-theoretic
identities, and states the present theorem boundaries for the probabilistic
and bootstrap constructions.

\begin{definition}[Classical Liouville datum]
\label{def:classical-liouville-datum}
A classical Liouville datum consists of an oriented smooth closed surface
\(\Sigma\), a Riemannian metric \(g\) on \(\Sigma\), a real scalar field
\(\phi\in C^\infty(\Sigma,\mathbb R)\), parameters
\[
  b>0,\qquad Q\in\mathbb R,\qquad \mu>0,
\]
and the Euclidean action
\begin{equation}
  S_L[\phi;g]
  =
  \frac1{4\pi}
  \int_\Sigma \dd^2x\,\sqrt g\,
  \left(
    g^{ab}\partial_a\phi\,\partial_b\phi
    +Q R_g\phi
    +4\pi\mu\,\ee^{2b\phi}
  \right).
  \label{eq:liouville-classical-action}
\end{equation}
Here \(R_g\) is the scalar curvature of \(g\).  The exponential term is a
function of the smooth classical field in this definition; in the quantum
theory it is a renormalized random measure or a normal-ordered local field,
depending on the construction framework.
\end{definition}

Varying a compactly supported smooth \(\delta\phi\) gives
\[
  \delta S_L
  =
  \frac1{4\pi}
  \int_\Sigma\sqrt g\,
  \left(
    -2\Delta_g\phi+Q R_g+8\pi b\mu\ee^{2b\phi}
  \right)\delta\phi .
\]
Thus the classical Euler--Lagrange equation is
\begin{equation}
  -\Delta_g\phi+\frac Q2 R_g+4\pi b\mu\ee^{2b\phi}=0.
  \label{eq:liouville-classical-eom}
\end{equation}
On a flat complex coordinate patch \(z\), the holomorphic stress tensor
representative is
\begin{equation}
  T(z)
  =
  -:\!(\partial\phi)^2\!:(z)+Q\,\partial^2\phi(z).
  \label{eq:liouville-stress-tensor}
\end{equation}
The normal ordering in \eqref{eq:liouville-stress-tensor} is a free-field
short-distance prescription.  It is used here as a perturbative coordinate on
the Virasoro representation data; the nonperturbative probabilistic
construction is stated later as a separate object.

\begin{convention}[Free-field normalization]
\label{conv:liouville-free-field-normalization}
In this chapter the local free field has two-point singularity
\[
  \phi(z)\phi(w)\sim -\frac12\log |z-w|^2,
  \qquad
  \partial\phi(z)\partial\phi(w)\sim -\frac1{2(z-w)^2}.
\]
The holomorphic exponential field is denoted
\[
  V_\alpha(z,\bar z)=\mathopen{:}\ee^{2\alpha\phi(z,\bar z)}\mathclose{:}.
\]
With this convention the standard Liouville relation is
\begin{equation}
  Q=b+b^{-1}.
  \label{eq:liouville-Q-b-relation}
\end{equation}
\end{convention}

\paragraph{Improved central charge and Liouville weights.}
Assume the free-field OPE convention
\ref{conv:liouville-free-field-normalization}.  The stress tensor
\eqref{eq:liouville-stress-tensor} has Virasoro central charge
\begin{equation}
  c_L=1+6Q^2.
  \label{eq:liouville-central-charge}
\end{equation}
The exponential \(V_\alpha\) is Virasoro primary in the free-field
coordinate with conformal weights
\begin{equation}
  h_\alpha=\bar h_\alpha=\alpha(Q-\alpha).
  \label{eq:liouville-exponential-weight}
\end{equation}

This is an OPE computation.  The stress tensor is \(T=T_0+T_Q\), where
\[
  T_0=-:\!(\partial\phi)^2\!:,\qquad T_Q=Q\partial^2\phi.
\]
The double contraction in \(T_0(z)T_0(w)\) gives the central term
\(\frac{1}{2}(z-w)^{-4}\), hence \(c=1\) for \(Q=0\).  The mixed terms have
at most third-order poles and contribute to the transformation of \(T\), not
to the central term.  Since
\[
  \partial^2\phi(z)\partial^2\phi(w)
  =
  \partial_z^2\partial_w^2
  \left(-\frac12\log(z-w)\right)
  \sim
  \frac{3}{(z-w)^4},
\]
the improvement contributes \(3Q^2(z-w)^{-4}\).  Comparing
\[
  T(z)T(w)\sim
  \frac{c/2}{(z-w)^4}
  +\frac{2T(w)}{(z-w)^2}
  +\frac{\partial T(w)}{z-w}
\]
gives \(c=1+6Q^2\).

For the exponential,
\[
  \partial\phi(z)V_\alpha(w,\bar w)
  \sim
  -\frac{\alpha}{z-w}V_\alpha(w,\bar w),
  \qquad
  \partial^2\phi(z)V_\alpha(w,\bar w)
  \sim
  \frac{\alpha}{(z-w)^2}V_\alpha(w,\bar w).
\]
The double contraction of \(-:\!(\partial\phi)^2\!:\) contributes
\(-\alpha^2(z-w)^{-2}V_\alpha\), and the improvement contributes
\(\alpha Q(z-w)^{-2}V_\alpha\).  The simple-pole term is
\((z-w)^{-1}\partial_wV_\alpha\).  Therefore \(V_\alpha\) has the weight
\(\alpha(Q-\alpha)\).

The interaction \(\mu\ee^{2b\phi}\) is marginal precisely when
\(b(Q-b)=1\), which gives \(Q=b+b^{-1}\).  Thus
\eqref{eq:liouville-Q-b-relation} is an operational normalization condition:
it is the condition that the exponential potential have weight \((1,1)\) in
the background-charge representation.

\section{Exponential Operators and the Probabilistic Object}

Liouville correlation functions are infinite-dimensional Euclidean QFT
objects.  The rigorous probabilistic construction separates the constant mode
from the mean-zero Gaussian free field and interprets the exponential
interaction as Gaussian multiplicative chaos.

\begin{definition}[Seiberg domain in physics normalization]
\label{def:liouville-seiberg-domain}
For insertions \(V_{\alpha_1}(x_1),\ldots,V_{\alpha_n}(x_n)\) on a closed
surface \(\Sigma\) with Euler characteristic \(\chi(\Sigma)\), the Seiberg
domain is the set of momenta satisfying
\begin{equation}
  \alpha_i<\frac Q2\quad\text{for every }i,
  \qquad
  2\sum_{i=1}^n\alpha_i>Q\,\chi(\Sigma).
  \label{eq:liouville-seiberg-bound}
\end{equation}
The first inequalities are local integrability conditions near the marked
points.  The second inequality is the negative-constant-mode integrability
condition.  On the sphere it reads \(\sum_i\alpha_i>Q\).
\end{definition}

The zero-mode check is elementary and fixes the sign conventions.  Write
\(\phi=c+\phi_0\), with \(\int_\Sigma\phi_0\,\mathrm{vol}_g=0\).  The
curvature term contributes \(\exp[-Q\chi(\Sigma)c]\), while the insertions
contribute \(\exp[2c\sum_i\alpha_i]\).  As \(c\to-\infty\), integrability
requires \(2\sum_i\alpha_i-Q\chi(\Sigma)>0\).  The positive \(c\) direction
is controlled by the Liouville potential when its multiplicative-chaos
measure is nonzero.

\begin{definition}[Gaussian multiplicative chaos input]
\label{def:liouville-gmc-input}
Let \(X_g\) be a mean-zero Gaussian free field on \((\Sigma,g)\), normalized
by a Green kernel with logarithmic singularity \(-\log d_g(x,y)\).  For
\(\gamma\in(0,2)\), let \(X_{g,\varepsilon}\) be a smooth mollification at
scale \(\varepsilon\).  The subcritical Gaussian multiplicative-chaos measure
is the probability-limit object
\begin{equation}
  M_{\gamma,g}(\dd x)
  =
  \lim_{\varepsilon\downarrow0}
  \varepsilon^{\gamma^2/2}
  \ee^{\gamma X_{g,\varepsilon}(x)}\,\mathrm{vol}_g(\dd x)
  \label{eq:liouville-gmc-measure}
\end{equation}
in the topology of weak convergence in probability against continuous test
functions.  In Liouville theory \(\gamma=2b\) after translating from the
physics field \(\phi\) of Convention
\ref{conv:liouville-free-field-normalization} to the probabilistic GFF
normalization.
\end{definition}

\begin{quotedtheorem}[Probabilistic Liouville correlation functions]
\label{qthm:gkrv-liouville-correlators}
The probabilistic Liouville constructions of
David--Kupiainen--Rhodes--Vargas and Guillarmou--Kupiainen--Rhodes--Vargas
define Liouville correlation functions by integrating the constant mode
against the Gaussian multiplicative-chaos functional, with insertions in the
Seiberg domain \eqref{eq:liouville-seiberg-bound}.  In the regimes covered by
these theorems the construction gives finite correlation functions,
conformal covariance with central charge \(1+6Q^2\), and meromorphic
continuation in the insertion momenta after the initial probabilistic domain
is established.  These theorems supply the measure-theoretic construction of
the closed-surface Liouville path integral; the full Segal sewing theory for
higher-genus boundary operations is recorded below as a separate boundary
problem.
\end{quotedtheorem}

\paragraph{Constant-mode reduction of the probabilistic construction.}
The theorem just quoted has a concrete core.  Let
\[
  \gamma=2b,\qquad
  s(\boldsymbol\alpha,\Sigma)
  =
  2\sum_{i=1}^n\alpha_i-Q\chi(\Sigma).
\]
After separating the constant mode, write the mean-zero field in the
probabilistic normalization as \(X_g\), with chaos parameter \(\gamma\).  The
local covariance used in
Convention~\ref{conv:liouville-free-field-normalization} is
\(-\frac12\log |z-w|^2=-\log |z-w|\), so it has the same logarithmic
coefficient as the probabilistic GFF.  The only notation change is that a
physics vertex \(V_{\alpha_i}=\mathopen{:}\ee^{2\alpha_i\phi}\mathclose{:}\)
is the probabilistic insertion with exponent
\[
  \alpha_i^{\mathrm{prob}}=2\alpha_i .
\]
The \(c\)-dependent part of an \(n\)-point function has the form
\[
  \int_{-\infty}^{\infty}
  \exp\!\left(
    s(\boldsymbol\alpha,\Sigma)c
    -\mu\,\ee^{\gamma c}
       Z_g(X_g;\boldsymbol x,\boldsymbol\alpha)
  \right)\dd c,
\]
where the random functional multiplying the exponential potential is
\begin{equation}
  Z_g(X_g;\boldsymbol x,\boldsymbol\alpha)
  =
  \int_\Sigma
  \prod_{i=1}^n
    \exp\!\bigl(\gamma \alpha_i^{\mathrm{prob}}\,G_g(y,x_i)\bigr)\,
  M_{\gamma,g}(\dd y).
  \label{eq:liouville-gmc-insertion-functional}
\end{equation}
The Green kernel \(G_g\) records the singularity of the GFF conditioned by
the vertex insertions; all coincident-point self-contractions are absorbed
into the renormalized exponential fields.  For \(s>0\), the elementary
change of variable \(t=\mu Z_g\ee^{\gamma c}\) gives
\begin{equation}
  \int_{\mathbb R}
  \ee^{s c-\mu Z_g\ee^{\gamma c}}\,\dd c
  =
  \frac1\gamma\,
  \mu^{-s/\gamma}\,
  \Gamma(s/\gamma)\,
  Z_g^{-s/\gamma}.
  \label{eq:liouville-zero-mode-gamma-integral}
\end{equation}
This identity is the source of the global Seiberg inequality
\(s(\boldsymbol\alpha,\Sigma)>0\).  The local inequalities
\(\alpha_i<Q/2\) are the corresponding local integrability conditions for
the random measure \eqref{eq:liouville-gmc-insertion-functional} near the
marked points.  Thus the probabilistic correlation function in the initial
domain is the expectation of the explicit negative moment
\[
  \mathbb E\!\left[
    Z_g(X_g;\boldsymbol x,\boldsymbol\alpha)^{-s/\gamma}
  \right],
\]
multiplied by the deterministic Green-function and gamma-function factors
fixed above.  The analytic theorem boundary supplies existence of these
negative moments, Weyl covariance of the renormalized expression, and
meromorphic continuation away from the initial domain.

\begin{definition}[Reflection relation]
\label{def:liouville-reflection-relation}
The Liouville spectrum is parametrized by
\[
  \alpha=\frac Q2+\ii P,\qquad P\in\mathbb R_{\ge0},
\]
with conformal weight \(Q^2/4+P^2\).  The reflection relation is the
identification, at the level of meromorphic correlation functions,
\begin{equation}
  V_\alpha=R_b(\alpha)\,V_{Q-\alpha},
  \label{eq:liouville-reflection-relation}
\end{equation}
where \(R_b(\alpha)\) is the meromorphic reflection amplitude fixed by the
two-point normalization and equivalently by the DOZZ three-point function.
Thus a single bulk primary label is represented by the reflection orbit
\(\{\alpha,Q-\alpha\}\), with the normalization encoded in \(R_b(\alpha)\).
\end{definition}

\begin{definition}[Asymptotic scattering basis]
\label{def:liouville-scattering-basis}
Let \(\phi_0\) be the constant mode of \(\phi\) on a long cylinder.  In the
weak-coupling end \(\phi_0\to-\infty\), the Liouville potential is negligible
and the Hilbert space is described by the background-charge free field with
oscillators.  A momentum-\(P\) incoming state has asymptotic wavefunction
\begin{equation}
  \Psi_P^{\mathrm{in}}(\phi_0,\{\phi_n\})
  \sim
  \left(
    e^{2\ii P\phi_0}
    +S(P)e^{-2\ii P\phi_0}
  \right)
  \exp\!\left(-\sum_{n\ge1}n\phi_n\phi_{-n}\right),
  \qquad P\ge0,
  \label{eq:liouville-asymptotic-wavefunction}
\end{equation}
where \(S(P)\) is the unitary reflection phase in the scattering
normalization.  The corresponding incoming primary is represented
asymptotically by
\begin{equation}
  V_P^{\mathrm{in}}
  \sim
  e^{(Q+2\ii P)\phi}
  +S(P)e^{(Q-2\ii P)\phi},
  \qquad
  h_P=\frac{Q^2}{4}+P^2.
  \label{eq:liouville-incoming-primary}
\end{equation}
The delta-normalized primary used in the continuous spectrum is
\(V_P=S(P)^{-1/2}V_P^{\mathrm{in}}\), with
\[
  \langle V_P(1)V_{P'}(0)\rangle=\pi\delta(P-P')
\]
after fixing the conventional measure.  This is the canonical-quantization
version of the reflection orbit \(\alpha=Q/2+\ii P\sim Q/2-\ii P\).
\end{definition}

\begin{definition}[Continuous conformal-block decomposition]
\label{def:liouville-continuous-blocks}
For four normalizable external Liouville primaries
\(\alpha_j=Q/2+\ii P_j\), the bootstrap ansatz on the sphere is a direct
integral over the same spectrum:
\begin{equation}
  \left\langle
    V_{P_1}(\infty)V_{P_2}(1)V_{P_3}(z)V_{P_4}(0)
  \right\rangle
  =
  \int_0^\infty
  C(P_1,P_2,P)C(P_3,P_4,P)
  \left|
    \mathcal F_{P}\!\left[
      \begin{matrix}P_1&P_2\\ P_4&P_3\end{matrix}
    \right](z)
  \right|^2\,dP .
  \label{eq:liouville-continuous-block-decomposition}
\end{equation}
Here \(\mathcal F_P\) is the Virasoro conformal block with internal weight
\(Q^2/4+P^2\).  Equation
\eqref{eq:liouville-continuous-block-decomposition} is a direct-integral
decomposition.  Its analytic content includes the choice of Plancherel
measure, the location of possible discrete residues after analytic
continuation, and the crossing kernel for the Virasoro blocks.
\end{definition}

\section{Virasoro Blocks from Gram Matrices}
\label{sec:liouville-virasoro-block-gram-matrices}

The symbol \(\mathcal F_P\) in
\eqref{eq:liouville-continuous-block-decomposition} hides a representation
theoretic construction.  The construction is local in the moduli variable
\(z\) and does not require the DOZZ formula.  It uses only Virasoro Ward
identities, the contravariant form on a Verma module, and the three-point
matrix elements of descendants.

\begin{definition}[Four-point Virasoro block coefficients]
\label{def:virasoro-block-coefficients}
Fix a central charge \(c\), an internal highest weight \(h\), and four
external weights \(h_1,h_2,h_3,h_4\).  Consider the chiral four-point
geometry
\[
  V_1(\infty)V_2(1)V_3(z)V_4(0),
\]
with the intermediate channel separating \((1,2)\) from \((3,4)\).  Assume
that the contravariant Gram matrix on the Verma module \(\mathcal V_{c,h}\)
is nondegenerate through the level under discussion.  Through level \(3\)
the normalized Virasoro block is written
\[
  \mathcal F_h^{(c)}
  \!\left[
    \begin{matrix}h_1&h_2\\ h_4&h_3\end{matrix}
  \right]\!(z)
  =
  z^{h-h_3-h_4}
  \left(1+f_1 z+f_2 z^2+f_3 z^3+O(z^4)\right).
\]
The assumption of nondegeneracy excludes \(h=0\) at level \(1\) and excludes
the zeros of the relevant Kac determinants at higher level.  At those
parameters one must quotient by null states, take a meromorphic limit, or
work with logarithmic extensions; the inverse Gram matrices below are not
literal inverses on the unreduced Verma module.
\end{definition}

\begin{lemma}[Level-one and level-two block coefficients]
\label{lem:liouville-virasoro-block-level-two}
In the notation of Definition~\ref{def:virasoro-block-coefficients}, set
\[
  a_1=h+h_3-h_4,\qquad b_1=h+h_2-h_1,
\]
\[
  a_2=h+2h_3-h_4,\qquad b_2=h+2h_2-h_1,
\]
and
\[
  a_{11}=a_1(a_1+1),\qquad b_{11}=b_1(b_1+1).
\]
Then
\begin{equation}
  f_1=\frac{b_1a_1}{2h}.
  \label{eq:liouville-virasoro-block-level-one}
\end{equation}
Let
\[
  G_2=
  \begin{pmatrix}
    4h+c/2 & 6h\\
    6h & 4h(2h+1)
  \end{pmatrix},
  \qquad
  \Delta_2=\det G_2
  =
  2h\{16h^2+2(c-5)h+c\}.
\]
With the ordered level-two basis
\((L_{-2}|h\rangle,L_{-1}^2|h\rangle)\), the second coefficient is
\begin{equation}
\begin{aligned}
  f_2
  =
  \frac1{\Delta_2}
  \bigl(&
    4h(2h+1)b_2a_2
    -6h(b_2a_{11}+b_{11}a_2)
  \\
  &\quad
    +(4h+c/2)b_{11}a_{11}
  \bigr).
\end{aligned}
  \label{eq:liouville-virasoro-block-level-two}
\end{equation}
For Liouville blocks one substitutes
\[
  c=1+6Q^2,\qquad h=\frac{Q^2}{4}+P^2,\qquad
  h_i=\alpha_i(Q-\alpha_i).
\]
For generic real \(P\) the denominators above are nonzero; the degenerate
values are precisely the places where a BPZ quotient changes the block
recursion.
\end{lemma}

\begin{proof}
Let \(\rho_{12}(Y)\) be the normalized three-point matrix element with the
descendant \(L_{-Y}|h\rangle\) in
\(\langle V_1(\infty)V_2(1)\cdot\rangle\), and let \(\rho_{34}(Y)\) be the
corresponding matrix element for the OPE of \(V_3(z)V_4(0)\).  The Ward
identity gives, for \(n=1,2\),
\[
  \rho_{12}(L_{-n})=h+n h_2-h_1,\qquad
  \rho_{34}(L_{-n})=h+n h_3-h_4.
\]
For the repeated \(L_{-1}\) descendant one differentiates the three-point
function
\[
  \langle V_1(\infty)V_2(1)V_h(x)\rangle
  =
  (1-x)^{-h_2-h+h_1}
\]
and its right-channel analogue.  This gives
\[
  \rho_{12}(L_{-1}^2)=b_1(b_1+1),\qquad
  \rho_{34}(L_{-1}^2)=a_1(a_1+1).
\]

The level-one Gram matrix is \((2h)\), so the projector onto the
intermediate module gives \eqref{eq:liouville-virasoro-block-level-one}.  At
level \(2\), the Virasoro commutation relation gives
\[
  \langle L_{-2}h,L_{-2}h\rangle=4h+c/2,\qquad
  \langle L_{-2}h,L_{-1}^2h\rangle=6h,
\]
\[
  \langle L_{-1}^2h,L_{-1}^2h\rangle=4h(2h+1).
\]
Therefore the level-two contribution is
\[
  f_2=
  \begin{pmatrix}b_2&b_{11}\end{pmatrix}
  G_2^{-1}
  \begin{pmatrix}a_2\\ a_{11}\end{pmatrix}.
\]
Inverting the displayed \(2\times2\) matrix gives
\[
  G_2^{-1}
  =
  \frac1{\Delta_2}
  \begin{pmatrix}
    4h(2h+1) & -6h\\
    -6h & 4h+c/2
  \end{pmatrix},
\]
which is exactly \eqref{eq:liouville-virasoro-block-level-two}.
\end{proof}

\begin{lemma}[Level-three block coefficient and determinant]
\label{lem:liouville-virasoro-block-level-three}
In the same four-point geometry, assume that the level-three Gram matrix is
nondegenerate.  Put
\[
  a_3=h+3h_3-h_4,\qquad b_3=h+3h_2-h_1,
\]
\[
  a_{21}=a_2(a_1+2),\quad b_{21}=b_2(b_1+2),
  \qquad
  a_{111}=a_1(a_1+1)(a_1+2),\quad
  b_{111}=b_1(b_1+1)(b_1+2).
\]
With ordered basis
\[
  \bigl(
    L_{-3}|h\rangle,\,
    L_{-2}L_{-1}|h\rangle,\,
    L_{-1}^3|h\rangle
  \bigr),
\]
the level-three Gram matrix is
\begin{equation}
  G_3=
  \begin{pmatrix}
    6h+2c & 10h & 24h\\
    10h & h(c+8h+8) & 12h(3h+1)\\
    24h & 12h(3h+1) & 24h(h+1)(2h+1)
  \end{pmatrix}.
  \label{eq:liouville-level-three-gram-matrix}
\end{equation}
Its determinant factors as
\begin{equation}
  \det G_3
  =
  48h^2
  \bigl(16h^2+2(c-5)h+c\bigr)
  \bigl(3h^2+(c-7)h+c+2\bigr).
  \label{eq:liouville-level-three-gram-determinant}
\end{equation}
Let
\[
  A_3=\begin{pmatrix}a_3\\ a_{21}\\ a_{111}\end{pmatrix},
  \qquad
  B_3=\begin{pmatrix}b_3\\ b_{21}\\ b_{111}\end{pmatrix}.
\]
Then the third block coefficient is the finite projector expression
\begin{equation}
  f_3=B_3^{\mathsf T}G_3^{-1}A_3.
  \label{eq:liouville-virasoro-block-level-three}
\end{equation}
Moreover, with \(h,h_i\) fixed and \(c\to\infty\),
\begin{equation}
  f_3
  =
  \frac{b_{111}a_{111}}{24h(h+1)(2h+1)}
  +O(c^{-1}),
  \label{eq:liouville-level-three-global-limit}
\end{equation}
which is the level-three coefficient of the global \(SL_2\) block.
\end{lemma}

\begin{proof}
The same Ward identities used at level two give the level-three descendant
matrix elements.  For the single mode,
\(\rho_{34}(L_{-3})=a_3\) and \(\rho_{12}(L_{-3})=b_3\).  For the mixed
descendant, the adjoint modes act in the order \(L_1L_2\).  Thus \(L_2\)
first sends the right three-point function to \(a_2 z^{h-h_3-h_4+2}\), and
then \(L_1\) gives \(a_2(a_1+2)\).  The left vector is identical with
\((h_3,h_4)\) replaced by \((h_2,h_1)\), giving \(b_2(b_1+2)\).  Three
derivatives of the three-point function give the \(L_{-1}^3\) entries
\(a_{111}\) and \(b_{111}\).

The entries of \(G_3\) follow by commuting the positive modes in
\[
  \langle h|L_Y L_{-Y'}|h\rangle
\]
to the right using
\[
  [L_m,L_n]=(m-n)L_{m+n}
  +\frac c{12}m(m^2-1)\delta_{m+n,0},
  \qquad
  L_n|h\rangle=0\quad(n>0),\quad L_0|h\rangle=h|h\rangle .
\]
For example,
\(\langle L_{-3}h,L_{-2}L_{-1}h\rangle
=\langle h|L_3L_{-2}L_{-1}|h\rangle=10h\), while
\(\langle L_{-2}L_{-1}h,L_{-1}^3h\rangle=12h(3h+1)\).  Explicitly, the six
independent entries are
\[
\begin{aligned}
  \langle h|L_3L_{-3}|h\rangle&=6h+2c,\\
  \langle h|L_3L_{-2}L_{-1}|h\rangle&=10h,\\
  \langle h|L_3L_{-1}^3|h\rangle&=24h,\\
  \langle h|L_1L_2L_{-2}L_{-1}|h\rangle&=h(c+8h+8),\\
  \langle h|L_1L_2L_{-1}^3|h\rangle&=12h(3h+1),\\
  \langle h|L_1^3L_{-1}^3|h\rangle&=24h(h+1)(2h+1).
\end{aligned}
\]
These identities give \eqref{eq:liouville-level-three-gram-matrix}.  Taking
the determinant of this displayed matrix gives
\eqref{eq:liouville-level-three-gram-determinant}; the first nontrivial
factor is the level-two Kac factor and the second is the new level-three
factor.

The projector onto level \(3\) in the intermediate Verma module is
\(G_3^{-1}\) in this basis, so contraction with the left and right
three-point descendant vectors gives
\eqref{eq:liouville-virasoro-block-level-three}.  Finally, in
\eqref{eq:liouville-level-three-gram-matrix} the first two basis vectors have
norms growing with \(c\), while the \(L_{-1}^3\) norm is independent of \(c\).
Equivalently, the \((3,3)\) entry of \(G_3^{-1}\) tends to
\([24h(h+1)(2h+1)]^{-1}\) and all inverse entries involving
\(L_{-3}\) or \(L_{-2}L_{-1}\) are \(O(c^{-1})\).  This proves
\eqref{eq:liouville-level-three-global-limit}.
\end{proof}

\section{Elliptic Nome and Recursion Coordinates}
\label{sec:liouville-elliptic-recursion-coordinate}

The \(z\)-series in Definition~\ref{def:virasoro-block-coefficients} is the
local OPE expansion near \(z=0\).  For crossing and numerical bootstrap
questions it is better to use the elliptic nome of the four-punctured sphere.
This change of coordinate is not a cosmetic acceleration trick: it separates
the universal geometry of the double cover from the representation-theoretic
recursion in the intermediate Virasoro module.

\begin{definition}[Elliptic nome of the four-punctured sphere]
\label{def:liouville-elliptic-nome}
Let
\[
  K(z)=\frac{\pi}{2}\,
  {}_2F_1\!\left(\frac12,\frac12;1;z\right)
\]
be the complete elliptic integral on the principal branch.  For
\(z\in\mathbb C\setminus((-\infty,0]\cup[1,\infty))\), define
\begin{equation}
  q(z)=\exp\!\left(-\pi\,\frac{K(1-z)}{K(z)}\right).
  \label{eq:liouville-elliptic-nome}
\end{equation}
Equivalently, if \(|q|\) is small,
\begin{equation}
  z=\lambda(q)
  =
  \frac{\vartheta_2(q)^4}{\vartheta_3(q)^4},
  \label{eq:liouville-modular-lambda}
\end{equation}
where
\[
  \vartheta_2(q)=2q^{1/4}\sum_{n\ge0}q^{n(n+1)},
  \qquad
  \vartheta_3(q)=1+2\sum_{n\ge1}q^{n^2}.
\]
The first terms are
\begin{equation}
  \lambda(q)=16q-128q^2+704q^3+O(q^4).
  \label{eq:liouville-lambda-q-expansion}
\end{equation}
\end{definition}

\begin{remark}[First conversion from \(z\)-coefficients to \(q\)-coefficients]
\label{prop:liouville-z-to-q-block-coefficients}
Let the normalized block be as in
Definition~\ref{def:virasoro-block-coefficients}, and set
\[
  a=h-h_3-h_4 .
\]
Write its elliptic-coordinate expansion before the universal theta-function
factorization as
\begin{equation}
  \mathcal F_h^{(c)}
  \!\left[
    \begin{matrix}h_1&h_2\\ h_4&h_3\end{matrix}
  \right]\!(\lambda(q))
  =
  (16q)^a
  \left(1+g_1q+g_2q^2+O(q^3)\right).
  \label{eq:liouville-raw-q-block-expansion}
\end{equation}
Then
\begin{align}
  g_1
  &=
  16f_1-8a,
  \label{eq:liouville-q-block-g1}\\
  g_2
  &=
  256f_2-128(1+a)f_1+32a^2+12a .
  \label{eq:liouville-q-block-g2}
\end{align}
Equation~\eqref{eq:liouville-lambda-q-expansion} gives
\[
  \lambda(q)=16q(1-8q+44q^2+O(q^3)).
\]
Therefore
\[
  \lambda(q)^a
  =
  (16q)^a
  \left(
    1-8aq+(32a^2+12a)q^2+O(q^3)
  \right).
\]
The nonleading part of the \(z\)-block contributes
\[
  1+f_1\lambda(q)+f_2\lambda(q)^2+O(q^3)
  =
  1+16f_1q+(-128f_1+256f_2)q^2+O(q^3).
\]
Multiplying these two displayed series gives
\eqref{eq:liouville-q-block-g1}--\eqref{eq:liouville-q-block-g2}.
\end{remark}

\begin{quotedtheorem}[Zamolodchikov elliptic recursion]
\label{qthm:zamolodchikov-elliptic-recursion}
Fix a central charge \(c=1+6Q^2\), external weights \(h_i\), and a nondegenerate
internal weight \(h\).  After multiplying
\(\mathcal F_h^{(c)}(z)\) by the universal powers of \(z\), \(1-z\), \(16q\),
and \(\vartheta_3(q)\) determined by \eqref{eq:liouville-modular-lambda}, one
obtains an elliptic block
\[
  H(c,h_i,h;q)=1+O(q)
\]
which is meromorphic as a function of \(h\) and satisfies the recursive
expansion
\begin{equation}
  H(c,h_i,h;q)
  =
  1+
  \sum_{m,n\ge1}
  \frac{(16q)^{mn}R_{m,n}(c,h_i)}
       {h-h_{m,n}(c)}
  H(c,h_i,h_{m,n}(c)+mn;q).
  \label{eq:zamolodchikov-elliptic-recursion}
\end{equation}
Here
\[
  h_{m,n}(c)
  =
  \frac{Q^2-(mb+n b^{-1})^2}{4}
\]
when \(Q=b+b^{-1}\), and the residues \(R_{m,n}\) are explicit products of
the inverse null-vector norm residue and the two chiral fusion polynomials for
the pairs \((h_1,h_2)\) and \((h_3,h_4)\).
\end{quotedtheorem}

The theorem boundary in
Theorem~\ref{qthm:zamolodchikov-elliptic-recursion} is deliberate.  The
monograph has derived the local \(z\)-coefficients and the first
coordinate-conversion coefficients above.  The full recursion also requires
normalizing null vectors and chiral vertex-operator matrix elements at every
degenerate level and proving the analytic meromorphy/convergence statements
that make the infinite recursive expression meaningful.  The residue products
must be inserted with the same external-leg and block normalizations as the
DOZZ constants; otherwise the recursion is a formal mnemonic rather than a
defined object.

\section{The DOZZ Formula and Shift Checks}

The exact three-point function is the Dorn--Otto--Zamolodchikov--Zamolodchikov
structure constant.  It is theorem-level input for this chapter, but the
special functions and convention checks are part of the manuscript.

\begin{definition}[The \(\Upsilon_b\) function]
\label{def:upsilon-b-function}
For \(b>0\) and \(Q=b+b^{-1}\), \(\Upsilon_b(x)\) is the entire function
characterized by self-duality \(\Upsilon_b(x)=\Upsilon_{1/b}(x)\), reflection
\(\Upsilon_b(x)=\Upsilon_b(Q-x)\), zeros at
\[
  x=-mb-nb^{-1}
  \quad\text{and}\quad
  x=Q+mb+nb^{-1},
  \qquad m,n\in\mathbb Z_{\ge0},
\]
and shift relations
\begin{align}
  \Upsilon_b(x+b)
  &=
  \gamma(bx)\,b^{1-2bx}\,\Upsilon_b(x),
  \label{eq:upsilon-b-shift-b}\\
  \Upsilon_b(x+b^{-1})
  &=
  \gamma(x/b)\,b^{2x/b-1}\,\Upsilon_b(x),
  \label{eq:upsilon-b-shift-binverse}
\end{align}
where \(\gamma(x)=\Gamma(x)/\Gamma(1-x)\).  These equations fix the
normalization up to the conventional value of \(\Upsilon'_b(0)\), which is
kept explicit in the three-point constant.
\end{definition}

\begin{quotedtheorem}[DOZZ three-point function]
\label{qthm:dozz-formula}
In the normalization of Convention
\ref{conv:liouville-free-field-normalization}, the sphere three-point
function has the form
\[
  \left\langle
    V_{\alpha_1}(z_1)V_{\alpha_2}(z_2)V_{\alpha_3}(z_3)
  \right\rangle_{\mathbb P^1}
  =
  \frac{
    C_b(\alpha_1,\alpha_2,\alpha_3)
  }{
    |z_{12}|^{2(h_1+h_2-h_3)}
    |z_{13}|^{2(h_1+h_3-h_2)}
    |z_{23}|^{2(h_2+h_3-h_1)}
  },
\]
with \(h_i=\alpha_i(Q-\alpha_i)\) and
\begin{equation}
\begin{aligned}
  C_b(\alpha_1,\alpha_2,\alpha_3)
  &=
  \left[
    \pi\mu\,\gamma(b^2)\,b^{2-2b^2}
  \right]^{(Q-\alpha_1-\alpha_2-\alpha_3)/b}
  \\
  &\quad\times
  \frac{
    \Upsilon'_b(0)
    \prod_{i=1}^3\Upsilon_b(2\alpha_i)
  }{
    \Upsilon_b(\alpha_1+\alpha_2+\alpha_3-Q)
    \Upsilon_b(\alpha_1+\alpha_2-\alpha_3)
    \Upsilon_b(\alpha_1+\alpha_3-\alpha_2)
    \Upsilon_b(\alpha_2+\alpha_3-\alpha_1)
  } .
\end{aligned}
  \label{eq:dozz-structure-constant}
\end{equation}
The formula was proposed by Dorn--Otto and by the Zamolodchikovs and was
proved in the probabilistic construction by Kupiainen--Rhodes--Vargas.  Its
role here is to fix the normalization whose reflection and shift consequences
are checked in the representation-theoretic calculations below.
\end{quotedtheorem}

\paragraph{Shift-equation uniqueness mechanism.}
The DOZZ theorem boundary combines a construction theorem with a bootstrap
uniqueness statement.  The local part of the uniqueness argument is visible
from the two degenerate fields developed later in this chapter.  Fusing
\(V_{-b/2}\) with a generic \(V_\alpha\) gives the two channels
\(\alpha-b/2\) and \(\alpha+b/2\); the BPZ equation and hypergeometric
connection matrix give the \(b\)-shift ratio
\[
  \frac{C_b(\alpha_1+b/2,\alpha_2,\alpha_3)}
       {C_b(\alpha_1-b/2,\alpha_2,\alpha_3)}
  =
  \mathcal R_b(\alpha_1;\alpha_2,\alpha_3)
\]
displayed explicitly in
\eqref{eq:liouville-dozz-b-shift-ratio}.  The dual degenerate insertion
\(V_{-1/(2b)}\) gives the same statement with \(b\) replaced by \(b^{-1}\).
Thus any meromorphic three-point constant in the same normalization obeys two
independent finite-difference equations:
\[
  C(\alpha_1+b,\alpha_2,\alpha_3)
  =
  A_b(\alpha_1;\alpha_2,\alpha_3)\,
  C(\alpha_1,\alpha_2,\alpha_3),
\]
\[
  C(\alpha_1+b^{-1},\alpha_2,\alpha_3)
  =
  A_{1/b}(\alpha_1;\alpha_2,\alpha_3)\,
  C(\alpha_1,\alpha_2,\alpha_3),
\]
with \(A_b\) and \(A_{1/b}\) fixed by the degenerate OPE coefficients and
connection matrices.  The ratio of two solutions is therefore invariant under
translations by \(b\) and \(b^{-1}\) in each momentum.  For irrational
\(b^2\), these translations generate a dense subgroup of the real line; with
the standard analyticity and growth hypotheses the ratio is constant.  The
reflection relation and the residue normalization at the screening poles fix
that constant.  The probabilistic theorem supplies an independently
constructed solution satisfying these properties, and the special-function
formula \eqref{eq:dozz-structure-constant} is the explicit representative.

The shift relations \eqref{eq:upsilon-b-shift-b}--\eqref{eq:upsilon-b-shift-binverse}
are the convention-sensitive part of \eqref{eq:dozz-structure-constant}.  They
are what make the degenerate-field shift equations compatible with the
normalization of the exponential interaction.  The companion script
\texttt{calculation-checks/liouville\_bpz\_checks.py} verifies the level-two
null-vector arithmetic used below.

\paragraph{Scattering-normalized DOZZ representative.}
Set \(\alpha_j=Q/2+\ii P_j\).  After absorbing the cosmological-constant
power and the two-point reflection factors into the delta-normalized fields
of Definition~\ref{def:liouville-scattering-basis}, a standard
scattering-normalized three-point representative is
\begin{equation}
\begin{aligned}
  C(P_1,P_2,P_3)
  &=
  \frac{\Upsilon'_b(0)}
       {\Upsilon_b\!\left(Q/2+\ii(P_1+P_2+P_3)\right)}
  \\
  &\quad\times
  \prod_{j=1}^3
  \frac{
    \left(\Upsilon_b(2\ii P_j)\Upsilon_b(-2\ii P_j)\right)^{1/2}
  }{
    \Upsilon_b\!\left(
      Q/2+\ii(P_1+P_2+P_3)-2\ii P_j
    \right)
  } .
\end{aligned}
  \label{eq:liouville-p-basis-dozz}
\end{equation}
Different square-root and reflection-amplitude conventions multiply
\eqref{eq:liouville-p-basis-dozz} by phase factors depending separately on
the external legs; the four-point integrand
\eqref{eq:liouville-continuous-block-decomposition} is invariant after the
same convention is used for both three-point constants and two-point
normalization.

Substitute \(\alpha_j=Q/2+\ii P_j\) into
\eqref{eq:dozz-structure-constant}.  The four denominator arguments become
\[
  \frac Q2+\ii(P_1+P_2+P_3),\qquad
  \frac Q2+\ii(P_1+P_2+P_3)-2\ii P_j,\quad j=1,2,3.
\]
The numerator factors are
\(\Upsilon_b(Q+2\ii P_j)=\Upsilon_b(-2\ii P_j)\) by the reflection identity
\(\Upsilon_b(x)=\Upsilon_b(Q-x)\).  Passing from \(V_{\alpha_j}\) to the
delta-normalized scattering fields contributes one square root of the
two-point reflection normalization per external leg, replacing the numerator
by
\(\left(\Upsilon_b(2\ii P_j)\Upsilon_b(-2\ii P_j)\right)^{1/2}\) up to the
leg phases just mentioned.  The remaining \(\mu\)-dependent power is a
zero-mode convention for the origin of \(\phi_0\), so it is absent from this
representative.

\paragraph{Pole normalization and the reflection phase.}
The scattering normalization also determines the phase \(S(P)\) in
Definition~\ref{def:liouville-scattering-basis}.  Let
\[
  \epsilon=\frac Q2+\ii(P_1+P_2+P_3).
\]
When \(\epsilon\to0\), the pole of
\(\Upsilon_b(\epsilon)^{-1}\) in
\eqref{eq:liouville-p-basis-dozz} gives
\begin{equation}
  C(P_1,P_2,P_3)
  \sim
  \frac1{\epsilon}
  \prod_{j=1}^3
  \left(
    \frac{\Upsilon_b(2\ii P_j)}
         {\Upsilon_b(-2\ii P_j)}
  \right)^{1/2}.
  \label{eq:liouville-dozz-pole-reflection-product}
\end{equation}
The same pole can be read in the weak-coupling region
\(\phi_0\ll0\) from the linear-dilaton zero-mode integral.  Introduce a
reference cutoff by integrating the zero mode over
\(\phi<\phi_{\mathrm{ref}}-\frac1{2b}\log\mu\).  Since the background charge
contributes \(\ee^{-2Q\phi}\) in the three-point sphere computation, the
asymptotic integral is
\[
  \int^{\phi_{\mathrm{ref}}-\frac1{2b}\log\mu}
  \dd\phi\,
  \ee^{-2Q\phi}
  \prod_{j=1}^3
  S(P_j)^{-1/2}\ee^{(Q+2\ii P_j)\phi}
  \sim
  \frac1\epsilon
  \left(\mu\ee^{2b\phi_{\mathrm{ref}}}\right)^{-\epsilon/b}
  \prod_{j=1}^3 S(P_j)^{-1/2}.
\]
Comparison with
\eqref{eq:liouville-dozz-pole-reflection-product}, for arbitrary triples
approaching the pole hyperplane, gives
\begin{equation}
  S(P)
  =
  \left(\mu\ee^{2b\phi_{\mathrm{ref}}}\right)^{-2\ii P/b}
  \frac{\Upsilon_b(-2\ii P)}{\Upsilon_b(2\ii P)} .
  \label{eq:liouville-reflection-phase-upsilon}
\end{equation}
Using the two \(\Upsilon_b\) shift relations reduces this to the gamma-ratio
form
\begin{equation}
  S(P)
  =
  \left(\mu\ee^{2b\phi_{\mathrm{ref}}} b^{2b^2-2}\right)^{-2\ii P/b}
  \frac{\gamma(2\ii P/b)}
       {b^2\,\gamma(-2\ii bP)} .
  \label{eq:liouville-reflection-phase-gamma}
\end{equation}
The factor containing \(\phi_{\mathrm{ref}}\) is the origin convention for the
zero mode; changing it multiplies \(V_P^{\mathrm{in}}\) by a momentum-dependent
phase and leaves the direct-integral four-point function unchanged when the
two-point normalization is transformed simultaneously.

\section{Boundary Liouville States}
\label{sec:liouville-boundary-states}

Boundary Liouville theory is the first place in this volume where the
Cardy-state intuition from rational CFT has to be rebuilt rather than merely
generalized by notation.  The closed spectrum is the continuous direct
integral over \(P\ge0\), the boundary states are distributional wavefunctions
on that spectrum, and the annulus kernels are Fourier or hyperbolic kernels
rather than entries of a finite modular \(S\)-matrix.  This section fixes the
normalization used later in the BCFT chapter; it does not assert a complete
Segal-style sewing construction.

\begin{definition}[Boundary Liouville datum]
\label{def:boundary-liouville-datum}
Let \((\Sigma,g)\) be an oriented smooth Riemannian surface with nonempty
smooth boundary.  Let \(n\) be the outward unit normal, \(K_g\) the geodesic
curvature of \(\partial\Sigma\) in the convention paired with \(n\), and
\(\dd s_g\) the induced line element.  A boundary Liouville datum consists of
\(\phi\in C^\infty(\Sigma,\mathbb R)\), \(b>0\), \(Q\in\mathbb R\),
\(\mu>0\), and a boundary cosmological constant \(\mu_B\in\mathbb R\), with
action
\begin{equation}
\begin{aligned}
  S_L^\partial[\phi;g,\mu_B]
  &=
  \frac1{4\pi}
  \int_\Sigma \dd^2x\,\sqrt g\,
  \left(
    g^{ab}\partial_a\phi\,\partial_b\phi
    +Q R_g\phi
    +4\pi\mu\,\ee^{2b\phi}
  \right)
  \\
  &\quad
  +\int_{\partial\Sigma}\dd s_g\,
  \left(
    \frac{Q}{2\pi}K_g\phi+\mu_B\ee^{b\phi}
  \right).
\end{aligned}
  \label{eq:boundary-liouville-action}
\end{equation}
The factor \(QK_g/(2\pi)\) is not optional: it is the boundary partner of
the bulk curvature coupling and is needed for conformal covariance under
Weyl rescaling on surfaces with boundary.
\end{definition}

\paragraph{Boundary Euler equation.}
For fixed metric \(g\), the first variation of
\eqref{eq:boundary-liouville-action} with arbitrary smooth
\(\delta\phi\) is stationary exactly when the bulk equation
\eqref{eq:liouville-classical-eom} holds and the boundary field satisfies
\begin{equation}
  \nabla_n\phi+QK_g+2\pi b\mu_B\ee^{b\phi}=0
  \qquad\text{on }\partial\Sigma .
  \label{eq:boundary-liouville-boundary-condition}
\end{equation}
On the upper half-plane with the Euclidean metric, \(K_g=0\) and the outward
normal is \(n=-\partial_y\), so
\(-\partial_y\phi+2\pi b\mu_B\ee^{b\phi}=0\).
The bulk variation is the one already displayed in
\eqref{eq:liouville-classical-eom}, except that integration by parts now
also gives the boundary term
\[
  \frac1{2\pi}\int_{\partial\Sigma}\dd s_g\,(\nabla_n\phi)\,\delta\phi .
\]
The variation of the boundary curvature coupling and of the boundary
exponential, with \(g\) fixed, is
\[
  \int_{\partial\Sigma}\dd s_g\,
  \left(\frac{Q}{2\pi}K_g+b\mu_B\ee^{b\phi}\right)\delta\phi .
\]
Since the boundary value of \(\delta\phi\) is arbitrary, the coefficient of
\(\delta\phi\) on \(\partial\Sigma\) must vanish.  Multiplying that boundary
coefficient by \(2\pi\) gives
\eqref{eq:boundary-liouville-boundary-condition}.

In the quantum conformal normalization \(Q=b+b^{-1}\), the boundary
exponential
\[
  B_\beta(x)=\mathopen{:}\ee^{\beta\phi(x)}\mathclose{:}
\]
has boundary scaling dimension
\begin{equation}
  h^\partial_\beta=\beta(Q-\beta)
  \label{eq:boundary-liouville-boundary-weight}
\end{equation}
in the doubled free-field coordinate.  Thus \(B_b\) has dimension \(1\).
This is the perturbative Ward-identity reason why the coupling \(\mu_B\) is
a conformal boundary coupling.  Existence of the noncompact interacting
boundary measure is a separate problem; the formulas below are bootstrap
boundary data whose sewing status is recorded explicitly.

\begin{definition}[FZZT parameter and wavefunction]
\label{def:fzzt-wavefunction}
Assume \(b>0\), \(Q=b+b^{-1}\), \(\mu>0\), and write
\[
  \gamma(x)=\frac{\Gamma(x)}{\Gamma(1-x)} .
\]
The FZZT boundary parameter \(s\in\mathbb C\) is related to the boundary
cosmological constant by
\begin{equation}
  \cosh^2(\pi b s)
  =
  \frac{\mu_B^2}{\mu}\sin(\pi b^2),
  \label{eq:fzzt-mub-s-relation}
\end{equation}
with the equality interpreted by analytic continuation when the displayed
right hand side is outside the real positive range.  For physical Liouville
momentum \(P\in\mathbb R\), set
\begin{equation}
  u_b(P)
  =
  \frac1{\ii P}
  \left(\pi\mu\,\gamma(b^2)\right)^{-\ii P/b}
  \Gamma(1+2\ii bP)\Gamma(1+2\ii P/b)
  \label{eq:fzzt-u-wavefunction}
\end{equation}
as a meromorphic distribution in \(P\), and
\begin{equation}
  U_s(P)=u_b(P)\cos(2\pi sP).
  \label{eq:fzzt-p-wavefunction}
\end{equation}
With the Ishibashi states \(\langle\!\langle P|\) normalized compatibly with
\eqref{eq:ishibashi-overlap-character}, the formal FZZT boundary state is
\begin{equation}
  \langle B_s|
  =
  \int_0^\infty \frac{\dd P}{\pi}\,
  U_s(P)\,\langle\!\langle P| .
  \label{eq:fzzt-boundary-state}
\end{equation}
\end{definition}

\begin{quotedtheorem}[FZZT one-point function]
\label{qthm:fzzt-one-point-function}
For generic \(b\) and generic boundary parameter \(s\), the disk one-point
function of the bulk primary \(V_\alpha=\mathopen{:}\ee^{2\alpha\phi}
\mathclose{:}\) on the upper half-plane has the form
\begin{equation}
  \langle V_\alpha(z,\bar z)\rangle_s
  =
  \frac{U_s(\alpha)}{|z-\bar z|^{2h_\alpha}},
  \qquad h_\alpha=\alpha(Q-\alpha),
  \label{eq:fzzt-disk-one-point-form}
\end{equation}
where
\begin{equation}
  U_s(\alpha)
  =
  \frac2b
  \left(\pi\mu\,\gamma(b^2)\right)^{(Q-2\alpha)/(2b)}
  \Gamma(2b\alpha-b^2)
  \Gamma\!\left(\frac{2\alpha}{b}-\frac1{b^2}-1\right)
  \cosh\!\left((2\alpha-Q)\pi s\right).
  \label{eq:fzzt-alpha-one-point}
\end{equation}
The theorem boundary is the boundary bootstrap: the degenerate
\(V_{-b/2}\) and \(V_{-1/(2b)}\) shift equations, the reflection relation,
and the annulus consistency checks determine the meromorphic solution under
the standard nonresonance and analytic-continuation hypotheses.
\end{quotedtheorem}

The conversion from \eqref{eq:fzzt-alpha-one-point} to the
momentum-space wavefunction in Definition~\ref{def:fzzt-wavefunction} is
elementary but convention sensitive.  If \(\alpha=Q/2+\ii P\), then
\[
  2b\alpha-b^2=1+2\ii bP,\qquad
  \frac{2\alpha}{b}-\frac1{b^2}-1=\frac{2\ii P}{b}.
\]
Using \(\Gamma(1+x)=x\Gamma(x)\), the prefactor \(2/b\) changes
\(\Gamma(2\ii P/b)\) into \(\Gamma(1+2\ii P/b)/(\ii P)\), while
\(\cosh(2\pi\ii sP)=\cos(2\pi sP)\).  Hence
\eqref{eq:fzzt-alpha-one-point} gives exactly
\eqref{eq:fzzt-p-wavefunction}.

\begin{definition}[ZZ boundary states as FZZT differences]
\label{def:zz-boundary-state}
For positive integers \(m,n\), define the two imaginary FZZT parameters
\[
  s_{m,n}^{+}=\ii\left(\frac{m}{b}+nb\right),
  \qquad
  s_{m,n}^{-}=\ii\left(\frac{m}{b}-nb\right).
\]
In the FZZT normalization above, the formal ZZ boundary state is the finite
difference
\begin{equation}
  \langle ZZ_{m,n}|
  =
  \langle B_{s_{m,n}^{+}}|
  -
  \langle B_{s_{m,n}^{-}}| .
  \label{eq:zz-as-fzzt-difference}
\end{equation}
\end{definition}

\begin{remark}[ZZ wavefunction in FZZT normalization]
\label{prop:zz-wavefunction-fzzt-normalization}
The boundary state \(\langle ZZ_{m,n}|\) has the momentum-space expansion
\begin{equation}
  \langle ZZ_{m,n}|
  =
  \int_0^\infty\frac{\dd P}{\pi}\,
  U_{m,n}^{ZZ}(P)\,\langle\!\langle P|,
  \label{eq:zz-boundary-state-wavefunction}
\end{equation}
where
\begin{equation}
  U_{m,n}^{ZZ}(P)
  =
  2u_b(P)\,
  \sinh\!\left(\frac{2\pi mP}{b}\right)
  \sinh(2\pi nbP).
  \label{eq:zz-wavefunction}
\end{equation}
Insert \(s_{m,n}^{\pm}\) into \(\cos(2\pi sP)\).  Since
\(\cos(\ii X)=\cosh X\),
\[
\begin{aligned}
  U_{s_{m,n}^{+}}(P)-U_{s_{m,n}^{-}}(P)
  &=
  u_b(P)
  \left[
    \cosh\!\left(2\pi P\left(\frac mb+nb\right)\right)
    -
    \cosh\!\left(2\pi P\left(\frac mb-nb\right)\right)
  \right]
  \\
  &=
  2u_b(P)\,
  \sinh\!\left(\frac{2\pi mP}{b}\right)\sinh(2\pi nbP),
\end{aligned}
\]
using \(\cosh(X+Y)-\cosh(X-Y)=2\sinh X\sinh Y\).  The same identity is
verified as an exact Laurent-polynomial equality in
\texttt{calculation-checks/bcft\_cardy\_checks.py}.
\end{remark}

The discrete labels \((m,n)\) are therefore not new rational-Cardy labels.
They are residues or finite differences inside a continuous boundary family.
Their annulus interpretation uses the same open--closed equality as
Chapter~\ref{ch:boundary-conformal-field-theory}, but with distributional
wavefunctions and continuous characters.  In annulus-normalized conventions
one often rewrites the degenerate \((m,n)\) wavefunction by dividing by the
basic \((1,1)\) hyperbolic factors,
\[
  \frac{\sinh(2\pi mP/b)\sinh(2\pi nbP)}
       {\sinh(2\pi P/b)\sinh(2\pi bP)} ,
\]
which makes the finite degenerate fusion sums visible.  The algebraic
hyperbolic identities are elementary; the analytic sewing theorem for the
resulting nonrational BCFT is not elementary and remains part of the open
problem below.

\section{Degenerate Insertions and BPZ Equations}

The simplest degenerate Liouville momentum is \(\alpha=-b/2\).  Its conformal
weight is
\[
  h_{-b/2}
  =
  -\frac12-\frac{3b^2}{4}.
\]

\paragraph{Level-two Liouville null vector.}
Let \(c=1+6Q^2\), \(Q=b+b^{-1}\), and
\[
  h=-\frac12-\frac{3b^2}{4}.
\]
In a highest-weight Virasoro module with highest vector \(|h\rangle\), the
level-two vector
\begin{equation}
  \left(L_{-1}^2+b^2L_{-2}\right)|h\rangle
  \label{eq:liouville-level-two-null-vector}
\end{equation}
is annihilated by \(L_1\) and \(L_2\).  Hence it is a null vector whenever the
module is quotiented to the degenerate representation carried by
\(V_{-b/2}\).
For a candidate \((L_{-1}^2+\kappa L_{-2})|h\rangle\), the Virasoro
commutators give
\[
  L_1L_{-1}^2|h\rangle=(4h+2)L_{-1}|h\rangle,
  \qquad
  L_1L_{-2}|h\rangle=3L_{-1}|h\rangle,
\]
and
\[
  L_2L_{-1}^2|h\rangle=6h|h\rangle,
  \qquad
  L_2L_{-2}|h\rangle=(4h+c/2)|h\rangle .
\]
The \(L_1\) equation requires
\(\kappa=-(4h+2)/3=b^2\).  Substituting
\(h=-1/2-3b^2/4\) and \(c=1+6(b+b^{-1})^2\) into the \(L_2\) equation gives
\[
  6h+b^2(4h+c/2)=0.
\]
Thus both positive modes annihilate \eqref{eq:liouville-level-two-null-vector}.

\paragraph{Dual level-two null vector and BPZ equation.}
Let \(c=1+6Q^2\), \(Q=b+b^{-1}\), and
\[
  h^\vee=-\frac12-\frac{3}{4b^2}.
\]
In a highest-weight Virasoro module with highest vector \(|h^\vee\rangle\),
the level-two vector
\begin{equation}
  \left(L_{-1}^2+b^{-2}L_{-2}\right)|h^\vee\rangle
  \label{eq:liouville-dual-level-two-null-vector}
\end{equation}
is annihilated by \(L_1\) and \(L_2\).  Therefore the degenerate insertion
\(V_{-1/(2b)}\) satisfies the dual BPZ equation
\begin{equation}
  \left[
    \partial_z^2
    +b^{-2}\sum_{i=1}^n
    \left(
      \frac{h_i}{(z-z_i)^2}
      +\frac1{z-z_i}\partial_{z_i}
    \right)
  \right]
  \left\langle
    V_{-1/(2b)}(z)\prod_{i=1}^n V_{\alpha_i}(z_i)
  \right\rangle
  =
  0 ,
  \label{eq:liouville-dual-bpz-equation}
\end{equation}
for separated-point correlators in any domain where the Ward identity and
the degenerate quotient are valid.
The Virasoro commutator calculation just displayed applies to every candidate
\((L_{-1}^2+\kappa L_{-2})|h\rangle\).  Substituting
\(h=h^\vee\) gives from the \(L_1\) equation
\[
  \kappa=-\frac{4h^\vee+2}{3}=b^{-2}.
\]
The remaining \(L_2\) coefficient is
\[
  6h^\vee+b^{-2}(4h^\vee+c/2).
\]
With \(c=1+6(b+b^{-1})^2\) and
\(h^\vee=-1/2-3/(4b^2)\), this coefficient is zero.  Hence
\eqref{eq:liouville-dual-level-two-null-vector} is a null vector in the
degenerate quotient.  Inserting
\((L_{-1}^2+b^{-2}L_{-2})V_{-1/(2b)}=0\) into the same stress-tensor Ward
identity used below gives \eqref{eq:liouville-dual-bpz-equation}.  The
restriction to separated points is the same restriction as for
\eqref{eq:liouville-bpz-equation}; collision limits require OPE
renormalization and meromorphic continuation.

\begin{proposition}[BPZ differential equation]
\label{prop:liouville-bpz-equation}
Let
\[
  F(z;z_1,\ldots,z_n)
  =
  \left\langle
    V_{-b/2}(z)\prod_{i=1}^n V_{\alpha_i}(z_i)
  \right\rangle
\]
be a separated-point correlation function in a domain where the Virasoro Ward
identity and the degenerate null-vector quotient are valid.  Then \(F\)
satisfies
\begin{equation}
  \left[
    \partial_z^2
    +b^2\sum_{i=1}^n
    \left(
      \frac{h_i}{(z-z_i)^2}
      +\frac1{z-z_i}\partial_{z_i}
    \right)
  \right]F=0,
  \qquad
  h_i=\alpha_i(Q-\alpha_i).
  \label{eq:liouville-bpz-equation}
\end{equation}
\end{proposition}

\begin{proof}
The proof uses only two inputs: the degenerate quotient in the Verma module of
\(V_{-b/2}\), and the separated-point Virasoro Ward identity.  The
state-field relation gives
\[
  \left\langle
    (L_{-1}^2V_{-b/2})(z)\prod_iV_{\alpha_i}(z_i)
  \right\rangle
  =
  \partial_z^2 F .
\]
For the \(L_{-2}\) insertion, write
\[
  (L_{-2}V_{-b/2})(z)
  =
  \oint_z\frac{\dd w}{2\pi\ii}\,
  \frac{T(w)V_{-b/2}(z)}{w-z}.
\]
The contour may be moved, inside the separated-point domain, to the contours
around the other insertions and the contour at infinity.  The latter vanishes
for the sphere correlator after the usual Ward identity falloff.  Around
\(z_i\),
\[
  T(w)V_{\alpha_i}(z_i)
  \sim
  \frac{h_i}{(w-z_i)^2}V_{\alpha_i}(z_i)
  +
  \frac{1}{w-z_i}\partial_{z_i}V_{\alpha_i}(z_i),
\]
and expanding \(1/(w-z)\) at \(w=z_i\) gives
\[
  \left\langle
    (L_{-2}V_{-b/2})(z)\prod_iV_{\alpha_i}(z_i)
  \right\rangle
  =
  \sum_i
  \left(
    \frac{h_i}{(z-z_i)^2}
      +\frac1{z-z_i}\partial_{z_i}
    \right)F.
\]
The null-vector equation in the degenerate quotient,
\[
  (L_{-1}^2+b^2L_{-2})V_{-b/2}=0,
\]
can now be inserted into the correlator.  Combining the two displayed
insertions gives \eqref{eq:liouville-bpz-equation}.  This is an identity for
separated-point correlators in the specified Ward-identity domain; coincident
limits require the OPE and contact-term renormalization rather than evaluation
of the differential equation on the collision diagonals.
\end{proof}

The BPZ equation constrains correlators recursively because fusing
\(V_{-b/2}\) with a generic \(V_\alpha\) has only two allowed Liouville
momenta,
\[
  V_{-b/2}V_\alpha
  \sim
  C_+(\alpha)V_{\alpha-b/2}
  +
  C_-(\alpha)V_{\alpha+b/2}.
\]
The dual degenerate field \(V_{-1/(2b)}\) similarly shifts a generic momentum
to \(\alpha\pm1/(2b)\) and supplies the \(b^{-1}\)-shift equation.  The two
degenerate equations are logically separate: the first uses the null vector
\eqref{eq:liouville-level-two-null-vector}, while the second uses
\eqref{eq:liouville-dual-level-two-null-vector}.  Treating the second as an
unstated symmetry argument would hide one of the hypotheses needed in the
standard bootstrap determination of the DOZZ constant.
Solving \eqref{eq:liouville-bpz-equation} near one puncture and matching it
to the same equation near another puncture gives shift equations for the
three-point structure constants.  The DOZZ constant is the meromorphic
solution with the reflection, shift, and analyticity properties stated above.

\begin{lemma}[Screening computation of the degenerate OPE coefficient]
\label{lem:liouville-degenerate-screening-coefficient}
In the free-field normalization of Convention
\ref{conv:liouville-free-field-normalization}, and with the interaction
normalization in \eqref{eq:liouville-classical-action}, the degenerate OPE is
normalized so that \(C_+(\alpha)=1\).  The one-screening coefficient in the
second channel is the meromorphic function
\begin{equation}
\begin{aligned}
  C_-(\alpha)
  &=
  -\pi\mu\,
  \gamma(1-2b\alpha)\,
  \gamma(1+b^2)\,
  \gamma(2b\alpha-1-b^2)
  \\
  &=
  -\pi\mu\,
  \frac{\gamma(2b\alpha-1-b^2)}
       {\gamma(2b\alpha)\gamma(-b^2)} ,
\end{aligned}
  \label{eq:liouville-coulomb-gas-cminus}
\end{equation}
where \(\gamma(x)=\Gamma(x)/\Gamma(1-x)\).  Equivalently,
\[
  C_-(\alpha)
  =
  \pi\mu\,b^4\gamma(b^2)\,
  \frac{\gamma(2b\alpha-1-b^2)}{\gamma(2b\alpha)} .
\]
The equalities are identities of meromorphic functions of \(b\) and
\(\alpha\); the intermediate Coulomb-gas integral is defined by analytic
continuation.
\end{lemma}

\begin{proof}
The coefficient \(C_+\) is fixed by the no-screening normal-ordered product:
the charges add as \(\alpha-b/2\), and the free contraction gives the
expected OPE power
\[
  V_{-b/2}(z)V_\alpha(0)
  =
  |z|^{2b\alpha}V_{\alpha-b/2}(0)+\cdots .
\]
For the second channel one expands the interaction
\(\exp[-\mu\int V_b]\) to first order.  The charge of
\(V_b(w)V_{-b/2}(z)V_\alpha(0)\) is \(\alpha+b/2\), so this is the unique
single-screening contribution to the channel \(V_{\alpha+b/2}\).  The free
contractions give
\[
  -\mu\int_{\mathbb C}\dd^2 w\,
  |z|^{2b\alpha}\,|w-z|^{2b^2}\,|w|^{-4b\alpha}\,
  V_{\alpha+b/2}(0).
\]
Set \(w=zu\).  The \(z\)-dependent factor becomes
\[
  |z|^{2+2b^2-2b\alpha}
  =
  |z|^{2(h_{\alpha+b/2}-h_\alpha-h_{-b/2})},
\]
which is exactly the required OPE power in the second channel.  Thus
\[
  C_-(\alpha)
  =
  -\mu
  \int_{\mathbb C}\dd^2u\,
  |u|^{-4b\alpha}|1-u|^{2b^2}.
\]
For complex exponents \(a,c\) in a convergence domain, the
Dotsenko--Fateev beta integral is
\begin{equation}
  \int_{\mathbb C}\dd^2u\,|u|^{2a}|1-u|^{2c}
  =
  \pi\,
  \gamma(1+a)\,
  \gamma(1+c)\,
  \gamma(-1-a-c).
  \label{eq:liouville-dotsenko-fateev-beta-integral}
\end{equation}
Both sides are meromorphic in \(a,c\), so
\eqref{eq:liouville-dotsenko-fateev-beta-integral} defines the continued
value when the physical exponents do not lie in an absolute-convergence
strip.  Substituting \(a=-2b\alpha\) and \(c=b^2\) gives the first line of
\eqref{eq:liouville-coulomb-gas-cminus}.  The remaining forms follow from
\(\gamma(1-x)=\gamma(x)^{-1}\),
\(\gamma(1+b^2)=\gamma(-b^2)^{-1}\), and
\(\gamma(1+b^2)=-b^4\gamma(b^2)\).
\end{proof}

\begin{lemma}[Dual screening computation]
\label{lem:liouville-dual-screening-coefficient}
Assume, as a Coulomb-gas representation of the \(b^{-1}\)-shift equation,
that the dual screening operator \(V_{1/b}\) is available with formal
coefficient \(\widetilde\mu\).  This is an additional bootstrap-screening
input, not an extra term in the classical action
\eqref{eq:liouville-classical-action}.  Normalize the no-dual-screening
channel in
\[
  V_{-1/(2b)}V_\alpha
  \sim
  \widetilde C_+(\alpha)V_{\alpha-1/(2b)}
  +
  \widetilde C_-(\alpha)V_{\alpha+1/(2b)}
\]
by \(\widetilde C_+(\alpha)=1\).  Then the one-dual-screening coefficient is
the meromorphic function
\begin{equation}
\begin{aligned}
  \widetilde C_-(\alpha)
  &=
  -\pi\widetilde\mu\,
  \gamma\!\left(1-\frac{2\alpha}{b}\right)\,
  \gamma\!\left(1+\frac1{b^2}\right)\,
  \gamma\!\left(\frac{2\alpha}{b}-1-\frac1{b^2}\right)
  \\
  &=
  \pi\widetilde\mu\,b^{-4}\gamma(b^{-2})\,
  \frac{
    \gamma\!\left(\frac{2\alpha}{b}-1-\frac1{b^2}\right)}
       {\gamma(2\alpha/b)} .
\end{aligned}
  \label{eq:liouville-dual-coulomb-gas-cminus}
\end{equation}
\end{lemma}

\begin{proof}
The no-dual-screening product has charge \(\alpha-1/(2b)\) and fixes
\(\widetilde C_+(\alpha)=1\).  The channel
\(\alpha+1/(2b)\) requires one insertion of the dual screening
\(V_{1/b}(w)\).  Its free contractions give
\[
  -\widetilde\mu\int_{\mathbb C}\dd^2 w\,
  |z|^{2\alpha/b}\,|w-z|^{2/b^2}\,|w|^{-4\alpha/b}\,
  V_{\alpha+1/(2b)}(0).
\]
After setting \(w=zu\), the \(z\)-dependent factor is
\[
  |z|^{2+2/b^2-2\alpha/b}
  =
  |z|^{2(h_{\alpha+1/(2b)}-h_\alpha-h_{-1/(2b)})},
\]
so the scaling power matches the dual OPE channel.  The remaining
Dotsenko--Fateev integral is
\[
  -\widetilde\mu
  \int_{\mathbb C}\dd^2u\,
  |u|^{-4\alpha/b}|1-u|^{2/b^2}.
\]
Applying \eqref{eq:liouville-dotsenko-fateev-beta-integral} with
\(a=-2\alpha/b\) and \(c=b^{-2}\) gives the first line of
\eqref{eq:liouville-dual-coulomb-gas-cminus}.  The second line follows from
\(\gamma(1-x)=\gamma(x)^{-1}\) and
\(\gamma(1+b^{-2})=-b^{-4}\gamma(b^{-2})\).  The formula is therefore the
precise \(b\leftrightarrow b^{-1}\) counterpart of
Lemma~\ref{lem:liouville-degenerate-screening-coefficient}, with
\(\widetilde\mu\) kept explicit because the relation between \(\mu\) and
\(\widetilde\mu\) is part of the full bootstrap normalization, not of this
local integral alone.
\end{proof}

\begin{proposition}[Degenerate BPZ blocks and the DOZZ \(b\)-shift]
\label{prop:liouville-degenerate-hypergeometric-solutions}
Fix three generic momenta \(\alpha_1,\alpha_2,\alpha_3\), put the degenerate
field \(V_{-b/2}\) at \(z\), and use global conformal covariance to set the
other insertions at \(0,1,\infty\).  After extracting the elementary powers
fixed by the OPE with \(V_{\alpha_1}(0)\), the two local solutions of
\eqref{eq:liouville-bpz-equation} near \(z=0\) are Gauss hypergeometric
functions.  In a normalization where the two fusion channels are
\(\alpha_1-b/2\) and \(\alpha_1+b/2\), one may write
\begin{align}
  \mathcal F_-(z)
  &=
  z^{b\alpha_1}(1-z)^{b\alpha_2}
  {}_2F_1(A,B;C;z),\\
  \mathcal F_+(z)
  &=
  z^{1+b^2-b\alpha_1}(1-z)^{b\alpha_2}
  {}_2F_1(A-C+1,B-C+1;2-C;z),
\end{align}
with
\[
  A=b(\alpha_1+\alpha_2+\alpha_3-Q-b/2),\quad
  B=b(\alpha_1+\alpha_2-\alpha_3-b/2),\quad
  C=1+b(2\alpha_1-Q).
\]
Analytic continuation from the \(z=0\) channel to the \(z=1\) channel gives
the corresponding gamma-function connection coefficients, hence the
degenerate shift equation
\begin{equation}
  \frac{
    C_b(\alpha_1+b/2,\alpha_2,\alpha_3)
  }{
    C_b(\alpha_1-b/2,\alpha_2,\alpha_3)
  }
  =
  \mathcal R_b(\alpha_1;\alpha_2,\alpha_3),
  \label{eq:liouville-degenerate-shift-ratio}
\end{equation}
where, in the DOZZ normalization of \eqref{eq:dozz-structure-constant},
\begin{equation}
\begin{aligned}
  \mathcal R_b(\alpha_1;\alpha_2,\alpha_3)
  &=
  \frac{
    \gamma(2b\alpha_1)\,
    \gamma(2b\alpha_1-b^2)\,
    \gamma\!\left(b(\alpha_2+\alpha_3-\alpha_1-b/2)\right)
  }{
    \pi\mu\,\gamma(b^2)\,b^4
  }
  \\
  &\quad\times
  \frac1{\gamma\!\left(b(\alpha_1+\alpha_2+\alpha_3-Q-b/2)\right)}
  \\
  &\quad\times
  \frac1{\gamma\!\left(b(\alpha_1+\alpha_2-\alpha_3-b/2)\right)}
  \frac1{\gamma\!\left(b(\alpha_1+\alpha_3-\alpha_2-b/2)\right)} .
\end{aligned}
  \label{eq:liouville-dozz-b-shift-ratio}
\end{equation}
The same function is obtained from the hypergeometric connection matrix and
the OPE normalization
\[
  V_{-b/2}V_{\alpha_1}
  \sim
  C_+(\alpha_1)V_{\alpha_1-b/2}
  +C_-(\alpha_1)V_{\alpha_1+b/2}.
\]
The connection matrix used here is the standard one for the Gauss equation:
\begin{align}
  {}_2F_1(A,B;C;z)
  &=
  \frac{\Gamma(C)\Gamma(C-A-B)}
       {\Gamma(C-A)\Gamma(C-B)}
  {}_2F_1(A,B;A+B-C+1;1-z)
  \notag\\
  &\quad+
  (1-z)^{C-A-B}
  \frac{\Gamma(C)\Gamma(A+B-C)}
       {\Gamma(A)\Gamma(B)}
  {}_2F_1(C-A,C-B;C-A-B+1;1-z).
  \label{eq:liouville-hypergeometric-connection-formula}
\end{align}
For generic momenta, none of the denominator gamma functions in
\eqref{eq:liouville-hypergeometric-connection-formula} is at a pole.  When a
parameter hits a pole, the equation is understood by meromorphic
continuation, and logarithmic resonances must be treated separately.
\end{proposition}

\begin{proof}
With insertions fixed at \(0,1,\infty\), equation
\eqref{eq:liouville-bpz-equation} has only regular singular points
\(\{0,1,\infty\}\).  The indicial equation at \(0\) has exponents equal to
the OPE weight differences
\[
  h_{\alpha_1-b/2}-h_{\alpha_1}-h_{-b/2}=b\alpha_1,
  \qquad
  h_{\alpha_1+b/2}-h_{\alpha_1}-h_{-b/2}=1+b^2-b\alpha_1 .
\]
After factoring these powers and the analogous elementary factor at \(1\),
the remaining second-order Fuchsian equation is the hypergeometric equation
with parameters \(A,B,C\) above.  The standard connection formula for
Gauss hypergeometric functions,
\eqref{eq:liouville-hypergeometric-connection-formula}, expresses the same
two-dimensional solution space in the \(z=1\) fusion basis.  The exponents
at \(z=1\) are
\[
  b\alpha_2,
  \qquad
  b\alpha_2+C-A-B=1+b^2-b\alpha_2,
\]
which are the two OPE exponents for fusing \(V_{-b/2}\) with
\(V_{\alpha_2}(1)\).  Crossing symmetry of the four-point function with one
degenerate insertion identifies the products of OPE coefficients in the two
bases, giving
\eqref{eq:liouville-degenerate-shift-ratio}.

It remains to compute the right hand side in the DOZZ normalization.  Let
\(K=\pi\mu\,\gamma(b^2)b^{2-2b^2}\).  In the ratio
\[
  \frac{
    C_b(\alpha_1+b/2,\alpha_2,\alpha_3)
  }{
    C_b(\alpha_1-b/2,\alpha_2,\alpha_3)
  },
\]
the cosmological-constant factor contributes \(K^{-1}\).  The numerator
\(\Upsilon_b(2\alpha_1)\) contributes
\[
  \frac{\Upsilon_b(2\alpha_1+b)}
       {\Upsilon_b(2\alpha_1-b)}
  =
  \gamma(2b\alpha_1)\,
  \gamma(2b\alpha_1-b^2)\,
  b^{2+2b^2-8b\alpha_1}.
\]
For the denominator factors one uses
\(\Upsilon_b(x)/\Upsilon_b(x+b)=
\gamma(bx)^{-1}b^{-1+2bx}\) for the first three terms and
\(\Upsilon_b(y+b)/\Upsilon_b(y)=\gamma(by)b^{1-2by}\) for the last term,
with
\[
\begin{gathered}
  x_1=\alpha_1+\alpha_2+\alpha_3-Q-b/2,\qquad
  x_2=\alpha_1+\alpha_2-\alpha_3-b/2,\\
  x_3=\alpha_1+\alpha_3-\alpha_2-b/2,\qquad
  y=\alpha_2+\alpha_3-\alpha_1-b/2 .
\end{gathered}
\]
The powers of \(b\) from these four denominator ratios and the numerator
combine to \(b^{-2b^2-2}\); multiplying by \(K^{-1}\) leaves the total
factor \(1/(\pi\mu\,\gamma(b^2)b^4)\).  The remaining gamma functions are
exactly \eqref{eq:liouville-dozz-b-shift-ratio}.
\end{proof}

\section{Sewing Status and Open Boundary}

The closed-surface probabilistic theory, the bootstrap formula, and the
Segal/Kontsevich--Segal sewing formulation require explicit comparison maps.

\begin{openproblem}[Liouville sewing and functorial closure]
\label{op:liouville-functorial-closure-volume-v}
Construct, in a single framework, the Hilbert or topological vector spaces
assigned to circles, the multilinear maps assigned to bordered Riemann
surfaces, the boundary state spaces, and the sewing maps whose closed-surface
matrix elements reproduce the probabilistic Liouville correlators and the
DOZZ/bootstrap conformal blocks.  The comparison must specify the topology on
continuous-spectrum direct integrals, the domain of unbounded vertex
operators, and the convergence of the gluing integrals over
\(\alpha=Q/2+\ii P\).  This is the Volume V form of
the Liouville and nonrational-CFT open problem in the
Kontsevich--Segal functorial-QFT chapter.
\end{openproblem}

Thus the status is deliberately layered.  The classical action and
background-charge Ward algebra are derived in the chapter.  The
measure-theoretic path integral and the DOZZ formula are quoted theorem
inputs with named proof boundaries.  Full functorial sewing remains an open
problem for the monograph's foundational program.
