\chapter{KMS States and Thermal Correlators}
\label{ch:kms-states-thermal-correlators}

\ConventionsRealTime

Thermal equilibrium is a property of a state together with a chosen time
evolution.  The KMS condition expresses this property as an analytic boundary
condition on correlation functions.  It agrees with the Gibbs trace in a
finite system, but it does not require the trace formula to exist.  This is
why it is the correct starting point for infinite-volume thermal QFT, where
the partition function per se is not an observable.

\paragraph{Statistical mechanics as a QFT absorption route.}
The statistical-mechanics material in Volumes~X and~XI is used as input for
QFT only after it has been translated into observables, regulators, limits,
and framework data.  The finite Gibbs trace below is therefore the first
rung of a route, not the template for every later thermal field theory.  The
route has five gates.

First, equilibrium data enter through a \(C^\ast\)-dynamical system and a
KMS state.  A finite Hamiltonian calculation may verify the strip boundary
condition, detailed balance, and fluctuation--dissipation signs, but the
QFT statement is the existence of a state on an infinite-volume observable
algebra with a declared time evolution.  This is the layer developed in the
present chapter and used by the Kubo analysis of
Chapter~\ref{ch:thermal-spectral-kubo-transport}.

Second, transport and hydrodynamics use equilibrium correlation functions
only after the order of limits and the retained slow variables are declared.
Kubo formulae, the Ward-identity hydrodynamic family of
Chapter~\ref{ch:thermal-hydrodynamics-from-ward-identities}, the
Schwinger--Keldysh hydrodynamic effective theory of
Chapter~\ref{ch:thermal-sk-hydrodynamic-effective-actions}, and the
long-time-tail analysis of Chapter~\ref{ch:thermal-hydrodynamic-fluctuations}
are linked by this datum: conserved densities and sources define the
observable map, while the low-frequency and long-wavelength limits remain
extra hypotheses or theorems in a specified model.

Third, nonequilibrium statistical mechanics contributes finite probability
identities and representation changes before it contributes continuum QFT.
The proof architecture of
Chapter~\ref{ch:thermal-nonequilibrium-steady-states-open-systems} separates
path-measure fluctuation relations, MSRJD and Doi--Peliti rewritings, tilted
generators, empirical-flow costs, GKSL limits, and Schwinger--Keldysh
influence kernels from the later scaling claim that produces a stochastic or
open-system field theory.  Dynamic-critical labels such as Models A--H live
at that scaling layer until a model-specific renormalization and observable
limit have been supplied.

Fourth, finite-locality and topological laboratories test mechanisms without
settling continuum existence.  Lieb--Robinson bounds, quasi-adiabatic
continuation, flux-torus Hall responses, toric-code sectors, and finite gauge
or TQFT boundary models give controlled finite-regulator identities for
locality, stability, and generalized-symmetry claims.  Their role in the QFT
spine is to expose which hypotheses a continuum theorem must retain, as in
the lattice and local-QFT routes of
Chapters~\ref{ch:lattice-reflection-positivity},
\ref{ch:constructive-wilson-lattice-gauge-theory}, and
\ref{ch:constructive-lattice-continuum-local-qft}.

Fifth, the constructive and numerical material of Volume~XI asks whether the
regulated families actually produce QFT data.  Cluster expansions, rigorous
RG, stochastic quantization, continuum scaling windows, Monte Carlo
evidence, transfer matrices, Hamiltonian truncation, tensor networks, and
neural variational states are admissible coordinates only when accompanied
by regulator families, renormalized observables, error budgets, and an
OS/Wightman/AQFT reconstruction or comparison target, as organized in
Chapters~\ref{ch:constructive-status-routes},
\ref{ch:constructive-continuum-limits-scaling-windows},
\ref{ch:constructive-rigorous-renormalization-group},
\ref{ch:constructive-stochastic-quantization-singular-spde}, and
\ref{ch:constructive-hamiltonian-truncation-dlcq-benchmarks}.

This route prevents a common compression.  A finite lattice identity, a
stochastic action, a transfer-matrix spectrum, or a tensor-network fit is not
yet a continuum QFT theorem.  To be used at that level it must name the
observable map, the thermodynamic or infinite-volume limit, the scaling
window and topology, the positivity and locality conditions, and the
framework in which the limiting object is reconstructed or compared.

\section{Dynamical Algebras and Analytic Elements}

\begin{definition}[\(C^\ast\)-dynamical system]
\label{def:cstar-dynamical-system}
A \(C^\ast\)-dynamical system is a pair \((\Obs,\alpha)\), where \(\Obs\) is
a \(C^\ast\)-algebra and
\[
  \alpha:\mathbb R\longrightarrow \operatorname{Aut}(\Obs),
  \qquad t\longmapsto \alpha_t,
\]
is a strongly continuous one-parameter group of \(\ast\)-automorphisms.  Thus
\(\alpha_{t+s}=\alpha_t\alpha_s\), \(\alpha_0=\id\), and for each
\(A\in\Obs\) the function \(t\mapsto\alpha_t(A)\) is norm-continuous.
\end{definition}

The algebra \(\Obs\) is the algebra of bounded observables.  Unbounded local
fields, stress tensors, and currents are used later through smeared
operators, affiliated operators, or distributional limits.  The KMS condition
is first stated on bounded observables because products and analytic
continuation are then unambiguous.

\begin{definition}[Analytic element]
\label{def:kms-analytic-element}
An element \(A\in\Obs\) is entire analytic for \(\alpha\) if the map
\(t\mapsto\alpha_t(A)\) extends to an entire \(\Obs\)-valued function
\[
  z\longmapsto\alpha_z(A)
\]
whose restriction to the real axis is the original orbit.  If \(w\in\mathbb R\),
then \(\alpha_w(A)\) is again entire analytic and its continuation is
\[
  z\longmapsto \alpha_{z+w}(A).
\]
A subalgebra \(\Obs_{\rm an}\subset\Obs\) is an analytic core if it is norm
dense, \(\alpha_t\)-invariant, and consists of entire analytic elements.
\end{definition}

\begin{lemma}[Existence of analytic elements by Gaussian smearing]
\label{lem:kms-gaussian-analytic-elements}
Let \((\Obs,\alpha)\) be a \(C^\ast\)-dynamical system.  For
\(A\in\Obs\) and \(n>0\), define
\begin{equation}
  A_n
  =
  \sqrt{\frac{n}{\pi}}
  \int_{\mathbb R}\dd t\,\ee^{-nt^2}\alpha_t(A),
  \label{eq:kms-gaussian-analytic-smearing}
\end{equation}
where the integral is the Bochner integral in the norm of \(\Obs\).  Then
\(A_n\) is entire analytic and \(A_n\to A\) in norm as \(n\to\infty\).  Hence
entire analytic elements form a norm-dense \(\ast\)-subalgebra.
\end{lemma}

\begin{proof}
Strong continuity implies that the integral in
\eqref{eq:kms-gaussian-analytic-smearing} exists and that
\[
  \|A_n-A\|
  \leq
  \sqrt{\frac{n}{\pi}}\int_{\mathbb R}\dd t\,\ee^{-nt^2}
  \|\alpha_t(A)-A\| .
\]
After the change of variables \(u=\sqrt n\,t\), the integrand is bounded by
\(\pi^{-1/2}\ee^{-u^2}2\|A\|\) and converges to zero for each fixed \(u\);
dominated convergence gives \(A_n\to A\).  For complex \(z\), define
\[
  \alpha_z(A_n)
  =
  \sqrt{\frac{n}{\pi}}
  \int_{\mathbb R}\dd t\,\ee^{-n(t-z)^2}\alpha_t(A).
\]
On every compact subset of the \(z\)-plane, the norm of the integrand is
bounded by \(\|A\|\exp[-n t^2+C(1+|t|)]\), which is integrable in \(t\).
Differentiation under the integral is therefore justified by dominated
convergence, so \(z\mapsto\alpha_z(A_n)\) is entire.  If \(A\) and \(B\) are
entire analytic, the functions
\[
  z\longmapsto \alpha_z(A)\alpha_z(B),
  \qquad
  z\longmapsto \alpha_{\bar z}(A)^\ast
\]
are respectively holomorphic and holomorphic as Banach-space-valued maps; on
the real axis they equal \(\alpha_t(AB)\) and \(\alpha_t(A^\ast)\).  The
identity theorem gives the analytic continuations of \(AB\) and \(A^\ast\).
For the adjoint statement, if \(\ell\) is any continuous linear functional on
\(\Obs\) and \(\ell^\#(X)=\overline{\ell(X^\ast)}\), then
\(\ell(\alpha_{\bar z}(A)^\ast)=\overline{\ell^\#(\alpha_{\bar z}(A))}\);
this is holomorphic in \(z\) because \(w\mapsto\ell^\#(\alpha_w(A))\) is
entire.  Finite linear combinations then give a dense \(\ast\)-subalgebra.
\end{proof}

\section{The KMS Boundary Condition}

\begin{definition}[KMS state]
\label{def:thermal-volume-kms-state}
Let \(\beta>0\).  A state \(\omega:\Obs\to\mathbb C\) is a
\(\beta\)-KMS state for \(\alpha\) if, for every \(A,B\in\Obs_{\rm an}\), the
function
\[
  t\longmapsto \omega\!\left(A\alpha_t(B)\right)
\]
is the lower boundary value of a function \(F_{A,B}(z)\) that is holomorphic
on the open strip
\[
  S_\beta=\{z\in\mathbb C:0<\operatorname{Im}z<\beta\},
\]
continuous and bounded on the closed strip, and satisfies
\begin{equation}
  F_{A,B}(t)=\omega\!\left(A\alpha_t(B)\right),
  \qquad
  F_{A,B}(t+\ii\beta)=\omega\!\left(\alpha_t(B)A\right)
  \label{eq:volume-x-kms-boundary-condition}
\end{equation}
for all \(t\in\mathbb R\).  The analytic core may be replaced by any
norm-dense analytic core; the resulting condition extends by continuity to
the usual KMS condition on \(\Obs\).
\end{definition}

The inverse temperature \(\beta\) has dimensions of time when \(\hbar=1\).
The time evolution \(\alpha_t\) is part of the datum.  Changing the time
evolution, for example by adding a chemical-potential generator, changes the
KMS condition.

Figure~\ref{fig:kms-strip} displays the analytic strip in the KMS condition:
the two boundary values differ by the time evolution through imaginary time
\(\ii\beta\).

\begin{figure}[H]
\centering
\begin{tikzpicture}[>=Stealth,scale=1.0]
  \draw[->] (-3.4,0) -- (3.8,0) node[right] {\(\operatorname{Re}z=t\)};
  \draw[->] (-3.2,-0.25) -- (-3.2,2.2) node[above]
    {\(\operatorname{Im}z\)};
  \draw[blue!60!black,thick] (-3,0) -- (3,0);
  \draw[blue!60!black,thick] (-3,1.55) -- (3,1.55);
  \fill[blue!8] (-3,0) rectangle (3,1.55);
  \draw[dashed] (-3,1.55) -- (3,1.55);
  \node[below] at (0,0) {\(\omega(A\alpha_t(B))\)};
  \node[above] at (0,1.55) {\(\omega(\alpha_t(B)A)\)};
  \node[left] at (-3.2,1.55) {\(\beta\)};
  \draw[->,thick] (2.25,0.2) -- (2.25,1.35);
  \node[right] at (2.25,0.78) {analytic continuation};
\end{tikzpicture}
\caption{The KMS condition is a strip-analytic boundary condition.  The lower
edge is the ordered correlator \(A\alpha_t(B)\); the upper edge is obtained by
moving \(B\) past the thermal state.}
\label{fig:kms-strip}
\end{figure}

\paragraph{Finite Gibbs traces as KMS states.}
Let \(\Hilb\) be finite dimensional, let \(H=H^\ast\), and let
\(\alpha_t(A)=\ee^{\ii Ht}A\ee^{-\ii Ht}\) on
\(\Obs=\operatorname{End}(\Hilb)\).  For
\begin{equation}
  \omega_\beta(A)
  =
  \frac{\operatorname{Tr}(\ee^{-\beta H}A)}
       {\operatorname{Tr}(\ee^{-\beta H})},
  \label{eq:finite-gibbs-state}
\end{equation}
\(\omega_\beta\) is a \(\beta\)-KMS state.  To see the strip function
explicitly, choose an orthonormal eigenbasis \(H\ket m=E_m\ket m\), and write
\(A_{mn}=\bra mA\ket n\), \(B_{mn}=\bra mB\ket n\).  Then
\begin{equation}
  F_{A,B}(z)
  =
  \frac{1}{Z_\beta}
  \sum_{m,n}
  \ee^{-\beta E_m}
  A_{mn}B_{nm}\,
  \ee^{\ii(E_n-E_m)z},
  \qquad
  Z_\beta=\sum_m\ee^{-\beta E_m},
  \label{eq:finite-kms-strip-function}
\end{equation}
is entire because the sum is finite.  At \(z=t\) it equals
\(\omega_\beta(A\alpha_t(B))\).  At \(z=t+\ii\beta\),
\[
  F_{A,B}(t+\ii\beta)
  =
  \frac{1}{Z_\beta}
  \sum_{m,n}
  \ee^{-\beta E_n}
  A_{mn}B_{nm}\,
  \ee^{\ii(E_n-E_m)t}.
\]
Relabeling \(m\leftrightarrow n\) gives
\[
  F_{A,B}(t+\ii\beta)
  =
  \frac{1}{Z_\beta}
  \sum_{m,n}
  \ee^{-\beta E_m}
  B_{mn}A_{nm}\,
  \ee^{\ii(E_m-E_n)t}
  =
  \omega_\beta(\alpha_t(B)A),
\]
which is the boundary condition
\eqref{eq:volume-x-kms-boundary-condition}.  This finite-dimensional
calculation is the model for the general KMS definition.  Infinite-volume
thermal states require the strip analyticity and boundary condition to be
verified in the chosen operator-algebraic or distributional framework.

\begin{remark}[Chemical potentials]
\label{rem:kms-chemical-potential-evolution}
If \(Q_A\) are commuting conserved charges and \(\mu_A\) are chemical
potentials, the finite trace
\[
  Z^{-1}\operatorname{Tr}\exp[-\beta(H-\mu_AQ_A)]\,(\cdot)
\]
is KMS for the automorphism generated by \(H-\mu_AQ_A\).  Equivalently, it is
KMS for the original Hamiltonian time evolution together with a simultaneous
imaginary-time internal-symmetry twist.  The two descriptions must not be
mixed without declaring which automorphism defines equilibrium.
\end{remark}

\section{Spectral Form of Detailed Balance}

In this section the Fourier transform of a tempered distribution \(G(t)\) is
\begin{equation}
  \widetilde G(\omega)
  =
  \int_{\mathbb R}\dd t\,\ee^{\ii\omega t}G(t).
  \label{eq:thermal-fourier-convention}
\end{equation}
With this convention, positive \(\omega\) is the energy absorbed by the
operator insertion in the finite-volume spectral formulas below.

Let \(A,B\) be observables whose time-dependent correlation functions have
tempered boundary values.  Define
\begin{equation}
  G_{AB}^{>}(t)=\omega(A(t)B),
  \qquad
  G_{AB}^{<}(t)=\omega(BA(t)),
  \qquad
  A(t)=\alpha_t(A).
  \label{eq:thermal-greater-lesser-definitions}
\end{equation}
The KMS condition implies, in the distributional boundary-value sense,
\begin{equation}
  G_{AB}^{>}(t-\ii\beta)=G_{AB}^{<}(t),
  \label{eq:kms-greater-lesser-boundary}
\end{equation}
or equivalently the frequency-space detailed-balance relation
\begin{equation}
  \widetilde G_{AB}^{<}(\omega)
  =
  \ee^{-\beta\omega}\widetilde G_{AB}^{>}(\omega).
  \label{eq:kms-detailed-balance}
\end{equation}

\paragraph{Finite-volume spectral detailed balance.}
In the same finite-dimensional setting, the distributions
\(\widetilde G^{>}_{AB}\) and \(\widetilde G^{<}_{AB}\) are finite sums of
delta functions and obey \eqref{eq:kms-detailed-balance}.  Using the
eigenbasis above,
\[
  G_{AB}^{>}(t)
  =
  \frac1{Z_\beta}
  \sum_{m,n}
  \ee^{-\beta E_m}A_{mn}B_{nm}
  \ee^{\ii(E_m-E_n)t}.
\]
Therefore
\begin{equation}
  \widetilde G_{AB}^{>}(\omega)
  =
  \frac{2\pi}{Z_\beta}
  \sum_{m,n}
  \ee^{-\beta E_m}A_{mn}B_{nm}
  \delta(\omega-E_n+E_m).
  \label{eq:finite-volume-greater-spectral}
\end{equation}
Similarly,
\begin{equation}
  \widetilde G_{AB}^{<}(\omega)
  =
  \frac{2\pi}{Z_\beta}
  \sum_{m,n}
  \ee^{-\beta E_n}A_{mn}B_{nm}
  \delta(\omega-E_n+E_m),
  \label{eq:finite-volume-lesser-spectral}
\end{equation}
where the second formula follows by relabeling the indices in
\(\omega(BA(t))\).  On the support of the displayed delta function,
\(\ee^{-\beta E_n}=\ee^{-\beta\omega}\ee^{-\beta E_m}\), proving
\eqref{eq:kms-detailed-balance}.

Define the thermal spectral distribution
\begin{equation}
  \rho_{AB}(\omega)
  =
  \widetilde G_{AB}^{>}(\omega)
  -
  \widetilde G_{AB}^{<}(\omega).
  \label{eq:thermal-spectral-density-definition}
\end{equation}
When \(1-\ee^{-\beta\omega}\) is invertible as a multiplier on the relevant
test-function class, detailed balance gives
\begin{equation}
  \widetilde G_{AB}^{>}(\omega)
  =
  \frac{\rho_{AB}(\omega)}{1-\ee^{-\beta\omega}},
  \qquad
  \widetilde G_{AB}^{<}(\omega)
  =
  \frac{\ee^{-\beta\omega}\rho_{AB}(\omega)}
       {1-\ee^{-\beta\omega}} .
  \label{eq:kms-greater-lesser-from-rho}
\end{equation}
For conserved densities, the \(\omega=0\) sector may contain delta functions
or derivatives of delta functions.  Those pieces have to be separated before
dividing by \(1-\ee^{-\beta\omega}\).  This is the same mathematical issue
that later appears as susceptibilities, hydrodynamic poles, and contact
terms.

\begin{remark}[Bosonic fluctuation--dissipation identity]
\label{rem:bosonic-fdt-kms}
Assume \(A=A^\ast\), \(B=B^\ast\), and no singular zero-frequency term in the
frequency window under consideration.  Define the symmetrized spectrum
\[
  \widetilde G^{\rm sym}_{AB}(\omega)
  =
  \frac12\left(
  \widetilde G_{AB}^{>}(\omega)+
  \widetilde G_{AB}^{<}(\omega)
  \right).
\]
Then
\begin{equation}
  \widetilde G^{\rm sym}_{AB}(\omega)
  =
  \frac12\coth\!\left(\frac{\beta\omega}{2}\right)\rho_{AB}(\omega).
  \label{eq:bosonic-fluctuation-dissipation}
\end{equation}
This identity follows directly from detailed balance.  Insert
\eqref{eq:kms-greater-lesser-from-rho} into the definition of
\(\widetilde G^{\rm sym}_{AB}\):
\[
  \widetilde G^{\rm sym}_{AB}
  =
  \frac12\,\frac{1+\ee^{-\beta\omega}}
  {1-\ee^{-\beta\omega}}\rho_{AB}
  =
  \frac12\coth\!\left(\frac{\beta\omega}{2}\right)\rho_{AB}.
\]
The stated zero-frequency hypothesis is precisely the condition that this
division is an ordinary distributional operation in the chosen window.
\end{remark}

For odd operators in a \(\mathbb Z_2\)-graded theory, the KMS boundary
condition includes the parity sign obtained by moving an odd operator around
the thermal circle.  The same proof gives
\[
  \widetilde G^{<}_{AB}(\omega)
  =
  -\ee^{-\beta\omega}\widetilde G^{>}_{AB}(\omega)
\]
for two odd operators in the convention where \(G^<\) is not itself graded by
an additional sign.  Equivalently, one may absorb the sign into the definition
of the fermionic lesser function.  The convention must be declared before
writing Fermi factors.

\section{Linear Response and Retarded Correlators}

Let \(B(t)\) be coupled to a compactly supported real source \(f(t)\) by
\begin{equation}
  H_f(t)=H-f(t)B(t)
  \label{eq:thermal-source-coupling}
\end{equation}
in a finite system, or by the corresponding time-dependent derivation on an
analytic algebra.  In the interaction picture,
\begin{equation}
  U_f(t,-\infty)
  =
  1+\ii\int_{-\infty}^{t}\dd s\,f(s)B(s)+O(f^2).
  \label{eq:linear-response-interaction-evolution}
\end{equation}
Thus
\[
  U_f(t,-\infty)^{-1}A(t)U_f(t,-\infty)
  =
  A(t)
  +\ii\int_{-\infty}^{t}\dd s\,f(s)[A(t),B(s)]
  +O(f^2).
\]
Stationarity of \(\omega\) gives
\begin{equation}
  \delta\omega(A(t))
  =
  \int_{-\infty}^{t}\dd s\,f(s)\,
  G^R_{AB}(t-s),
  \qquad
  G^R_{AB}(t)
  =
  -\ii\,\theta(t)\,\omega([A(t),B(0)]) .
  \label{eq:thermal-linear-response-retarded}
\end{equation}
The retarded distribution is causal: its support in the time variable lies in
\(t\geq0\).  Consequently, under the usual polynomial-growth hypotheses for
tempered distributions, its Fourier transform is the boundary value from the
upper half \(\omega\)-plane.

Figure~\ref{fig:retarded-support} records the causal support property of the
retarded response function, which is the input for its upper-half-plane
analytic Fourier transform.

\begin{figure}[H]
\centering
\begin{tikzpicture}[>=Stealth,scale=0.95]
  \draw[->] (-3.5,0) -- (3.7,0) node[right] {\(t\)};
  \draw[very thick,blue!65!black] (0,0) -- (3.2,0);
  \draw[very thick,gray!45] (-3.2,0) -- (0,0);
  \draw[fill=blue!65!black] (0,0) circle (0.045);
  \node[below] at (-1.8,0) {no response};
  \node[below] at (1.65,0) {retarded support};
  \draw[->,thick] (0,0.3) -- (0,1.0);
  \node[above] at (0,1.0) {source insertion};
\end{tikzpicture}
\caption{The retarded kernel is supported only after the source insertion.
This causal support, not Euclidean periodicity alone, is what gives the
upper-half-plane analyticity of \(G^R(\omega)\).}
\label{fig:retarded-support}
\end{figure}

For translation-invariant thermal QFT, the retarded two-point function of
local densities is
\begin{equation}
  G^R_{AB}(\omega,\mathbf k)
  =
  -\ii\int_0^\infty \dd t\int \dd^{D-1}x\,
  \ee^{\ii\omega t-\ii\mathbf k\cdot\mathbf x}
  \omega([A(t,\mathbf x),B(0,\mathbf0)]) .
  \label{eq:thermal-retarded-qft-definition}
\end{equation}
With the Fourier convention \eqref{eq:thermal-fourier-convention}, the
spectral function associated to \(G^R\) is
\begin{equation}
  \rho_{AB}(\omega,\mathbf k)
  =
  -2\,\operatorname{Im}G^R_{AB}(\omega+\ii0,\mathbf k).
  \label{eq:thermal-rho-retarded-sign}
\end{equation}
For Hermitian \(A=B\), positivity of the finite-volume spectral weights gives
\(\rho_{AA}(\omega,\mathbf k)\geq0\) for \(\omega>0\) before the thermodynamic
limit.  The infinite-volume statement is obtained by distributional limits,
provided the local algebra and state have been constructed.

\section{Conserved Densities and Hydrodynamic Variables}

Hydrodynamics begins with conserved local currents in the microscopic theory.
Let \(T^{\mu\nu}\) be the stress tensor and let \(J_A^\mu\) be conserved
currents for ordinary internal symmetries:
\begin{equation}
  \partial_\mu T^{\mu\nu}=0,
  \qquad
  \partial_\mu J_A^\mu=0 .
  \label{eq:thermal-conservation-laws}
\end{equation}
In a homogeneous isotropic thermal state with unit future-timelike vector
\(u^\mu\), temperature \(T=\beta^{-1}\), and chemical potentials \(\mu_A\),
the one-point functions define thermodynamic densities by
\begin{equation}
  \omega(T^{\mu\nu})
  =
  \varepsilon u^\mu u^\nu+p\Delta^{\mu\nu},
  \qquad
  \omega(J_A^\mu)=n_A u^\mu,
  \qquad
  \Delta^{\mu\nu}=\eta^{\mu\nu}+u^\mu u^\nu .
  \label{eq:thermal-homogeneous-one-point-functions}
\end{equation}
Here \(\varepsilon\) is the energy density, \(p\) the pressure, and \(n_A\) the
charge densities.  These are numbers attached to a homogeneous equilibrium
state.  They become hydrodynamic fields only after one assumes a family of
near-equilibrium states whose long-wavelength connected correlators are
controlled by slowly varying local values of
\[
  T(x),\qquad \mu_A(x),\qquad u^\mu(x).
\]

The derivative expansion is a statement about constitutive relations:
\begin{equation}
  T^{\mu\nu}
  =
  T^{\mu\nu}_{(0)}+T^{\mu\nu}_{(1)}+T^{\mu\nu}_{(2)}+\cdots,
  \qquad
  J_A^\mu
  =
  J_{A,(0)}^\mu+J_{A,(1)}^\mu+J_{A,(2)}^\mu+\cdots,
  \label{eq:thermal-hydrodynamic-derivative-expansion}
\end{equation}
where the subscript records the number of derivatives acting on
\(T,\mu_A,u^\mu\) and on background sources.  Existence of such an expansion
requires hypotheses beyond the KMS boundary condition: it is a long-time,
long-distance property of the QFT state.

\paragraph{First hydrodynamic poles from conservation.}
Let a homogeneous parity-even thermal state in \(D=d_s+1\) spacetime
dimensions have enthalpy \(w=\varepsilon+p>0\), speed of sound
\[
  c_s^2=\left(\frac{\partial p}{\partial\varepsilon}\right)_{\!n_A/s},
\]
shear viscosity \(\eta_{\rm sh}\), bulk viscosity \(\zeta\), and a decoupled
charge-diffusion channel with conductivity \(\sigma\) and susceptibility
\(\chi>0\).  Linearizing the first-order constitutive relations around the
homogeneous state gives the following small-\(|\mathbf k|\) hydrodynamic
denominators, and therefore the corresponding retarded poles in every
conserved-current correlator whose residue on the mode is nonzero:
\begin{equation}
  \omega_{\rm shear}(\mathbf k)
  =
  -\,\ii\,\frac{\eta_{\rm sh}}{w}\,\mathbf k^2+O(|\mathbf k|^4),
  \label{eq:kms-shear-pole}
\end{equation}
\begin{equation}
  \omega_{\rm charge}(\mathbf k)
  =
  -\,\ii\,\frac{\sigma}{\chi}\,\mathbf k^2+O(|\mathbf k|^4),
  \label{eq:kms-charge-diffusion-pole}
\end{equation}
and
\begin{equation}
  \omega_{\rm sound}^{\pm}(\mathbf k)
  =
  \pm c_s|\mathbf k|
  -\frac{\ii}{2w}
  \left(\zeta+\frac{2(d_s-1)}{d_s}\eta_{\rm sh}\right)
  \mathbf k^2
  +O(|\mathbf k|^3).
  \label{eq:kms-sound-poles}
\end{equation}

Take \(\mathbf k\) in the \(x\)-direction and use the Fourier convention
\(\ee^{-\ii\omega t+\ii kx}\).  In the shear channel a transverse velocity
perturbation \(v_y\) obeys
\[
  -\ii\omega\,w\,v_y+\eta_{\rm sh}k^2v_y=0,
\]
which gives \eqref{eq:kms-shear-pole}.  In a decoupled charge channel,
Fick's law \(J_A^x=-\sigma\,\partial_x(\delta\mu_A)\) and
\(\delta n_A=\chi\,\delta\mu_A\) give
\[
  -\ii\omega\,\delta n_A+\frac{\sigma}{\chi}k^2\delta n_A=0,
\]
which gives \eqref{eq:kms-charge-diffusion-pole}.  For longitudinal sound,
the linearized conservation equations are
\[
  -\ii\omega\,\delta\varepsilon+\ii k\,w\,v_x=0,
\]
\[
  -\ii\omega\,w\,v_x+\ii k\,c_s^2\delta\varepsilon
  +
  \left(\zeta+\frac{2(d_s-1)}{d_s}\eta_{\rm sh}\right)k^2v_x=0 .
\]
Eliminating \(v_x\) gives
\[
  \omega^2-c_s^2k^2
  +
  \frac{\ii}{w}
  \left(\zeta+\frac{2(d_s-1)}{d_s}\eta_{\rm sh}\right)
  \omega k^2=0 .
\]
Solving this quadratic at small \(k\) gives \eqref{eq:kms-sound-poles}.
This calculation identifies the hydrodynamic poles of the retarded functions
once the linearized constitutive relations, the long-distance derivative
expansion, and the existence of the corresponding retarded correlators have
been established.

Figure~\ref{fig:thermal-hydrodynamic-variables} depicts this analytic
content of the hydrodynamic limit.  The hydrodynamic variables are not new
microscopic operators; they are the coordinates whose linearized equations
describe the poles of retarded conserved-current correlators that approach
the origin as \(|\mathbf k|\to0\).

\begin{figure}[H]
\centering
\begin{tikzpicture}[>=Stealth,scale=1.0]
  \draw[->] (-3.4,0) -- (3.6,0) node[right] {\(\operatorname{Re}\omega\)};
  \draw[->] (0,-3.0) -- (0,1.05) node[above] {\(\operatorname{Im}\omega\)};
  \fill[blue!70!black] (-1.75,-0.38) circle (2.1pt);
  \fill[blue!70!black] (1.75,-0.38) circle (2.1pt);
  \node[blue!70!black,above left] at (-1.75,-0.38)
    {\(\omega^-_{\rm sound}\)};
  \node[blue!70!black,above right] at (1.75,-0.38)
    {\(\omega^+_{\rm sound}\)};
  \fill[green!45!black] (0,-0.65) circle (2.1pt);
  \node[green!45!black,right] at (0.08,-0.65)
    {\(-\ii D_\eta\mathbf k^2\)};
  \fill[orange!85!black] (0,-1.0) circle (2.1pt);
  \node[orange!85!black,right] at (0.08,-1.0)
    {\(-\ii D_c\mathbf k^2\)};
  \draw[dashed] (-3.1,-1.55) -- (3.1,-1.55);
  \node[anchor=west] at (0.15,-1.78)
    {nonhydrodynamic gap};
  \fill[red!70!black] (-1.45,-2.15) circle (2pt);
  \fill[red!70!black] (1.25,-2.35) circle (2pt);
  \fill[red!70!black] (0.35,-2.55) circle (2pt);
  \node[red!70!black,below] at (0.0,-2.75)
    {microscopic relaxation poles};
  \draw[decorate,decoration={brace,amplitude=4pt}]
    (2.85,-1.55) -- (2.85,0);
  \node[right] at (2.95,-0.78) {\(O(\tau_{\rm micro}^{-1})\)};
  \node[below left] at (0,0) {\(0\)};
  \node[anchor=west] at (0.35,0.62)
    {\(\operatorname{Im}\omega>0\): analytic};
\end{tikzpicture}
\caption{At fixed small \(|\mathbf k|\), hydrodynamics appears in retarded
thermal correlators as poles that approach the origin: two sound poles, a
shear-diffusion pole with \(D_\eta=\eta_{\rm sh}/w\), and charge-diffusion
poles with \(D_c=\sigma/\chi\) in a decoupled channel.  Nonhydrodynamic
relaxation poles remain separated by a microscopic gap in a regime where the
derivative expansion is controlled.}
\label{fig:thermal-hydrodynamic-variables}
\end{figure}

At first order, parity-even relativistic hydrodynamics in flat spacetime
contains shear viscosity \(\eta_{\rm sh}\), bulk viscosity \(\zeta\), and
charge-conductivity data.  In the shear channel, couple a metric perturbation
\(h_{xy}(t)\) to \(T^{xy}\).  The linear-response formula gives
\[
  \delta\omega(T^{xy}(\omega))
  =
  -\frac12G^R_{T^{xy}T^{xy}}(\omega,\mathbf0)h_{xy}(\omega)
  \]
with the factor \(1/2\) inherited from the metric coupling
\(\frac12\int T^{ij}h_{ij}\).  Comparing with the constitutive relation fixes
\begin{equation}
  \eta_{\rm sh}
  =
  \lim_{\omega\downarrow0}
  \frac{\rho_{T^{xy}T^{xy}}(\omega,\mathbf0)}{2\omega}
  =
  -\lim_{\omega\downarrow0}
  \frac{\operatorname{Im}G^R_{T^{xy}T^{xy}}(\omega,\mathbf0)}{\omega}.
  \label{eq:thermal-shear-kubo}
\end{equation}
This is a definition of the transport coefficient in terms of QFT correlators
once the zero-frequency limit exists.  Later chapters derive the
constitutive relations, entropy-current constraints, Schwinger--Keldysh
effective action, anomaly-induced transport, and kinetic limits from these
correlator-level objects.

\begin{openproblem}[Thermal real-time reconstruction]
\label{op:thermal-real-time-reconstruction}
Give constructive hypotheses under which Euclidean finite-temperature
Schwinger functions on \(S^1_\beta\times\mathbb R^{D-1}\) reconstruct a
real-time KMS QFT with local observables, spectral functions, and transport
coefficients.  The required analytic continuation is more delicate than the
vacuum OS reconstruction because thermal correlators have periodic Euclidean
time and no vacuum spectrum condition.
\end{openproblem}
