\chapter{Nonequilibrium Steady States And Open-System Limits}
\label{ch:thermal-nonequilibrium-steady-states-open-systems}

\ConventionsRealTime

Nonequilibrium steady states and open systems are real-time QFT objects.  A
steady state is a positive invariant state on an algebra of observables.  An
open-system limit is a scaling limit of a larger closed system in which a
subalgebra evolves by a completely positive trace-preserving semigroup.  This
chapter fixes these definitions, derives the reservoir entropy-production
formula, records finite path-measure fluctuation identities before taking a
continuum limit, derives the finite-step Langevin/MSRJD representation,
derives the finite reaction-network Doi--Peliti representation, records the
tilted-generator construction of dynamical large deviations, records the
weak-coupling route to the GKSL generator, and connects the Markovian limit
to a Schwinger--Keldysh influence functional and to hydrodynamic relaxation
with noise.

\section{Steady-State Datum}

Let \(\mathcal A\) be a \(*\)-algebra of observables for a real-time QFT,
or a local \(C^*\)-algebra when a norm-completed formulation is available.
A state is a positive normalized linear functional
\[
  \omega:\mathcal A\to\mathbb C,
  \qquad
  \omega(A^*A)\ge0,\qquad
  \omega(\mathbf 1)=1.
\]
Let \(\tau_t\) be the time-evolution automorphism group.  A steady state is a
state \(\omega_{\rm ss}\) such that
\begin{equation}
  \omega_{\rm ss}(\tau_t(A))=\omega_{\rm ss}(A)
  \label{eq:ness-invariant-state}
\end{equation}
for all \(A\) in the common domain of the state and the time evolution.
Local stationarity means that the distributional correlation functions of
local observables are invariant under simultaneous time translation.  A
nonequilibrium steady state is a steady state whose current expectation
values or reservoir boundary data are not those of a single KMS state.

Equation \eqref{eq:ness-invariant-state} is only the kinematic definition.
Existence, uniqueness, clustering, stability under changes of the preparation
protocol, and local normality relative to reference states are separate
properties.  They must be proved or assumed in any concrete model.

\section{Reservoir Construction}

Consider a QFT on a line prepared at \(t=0\) by joining two half-systems.
The left and right initial states are KMS states with parameters
\[
  (\beta_L,\mu_L),
  \qquad
  (\beta_R,\mu_R).
\]
For a local observable \(A\) supported near the junction, a central steady
state is a weak limit
\begin{equation}
  \omega_{\rm ss}(A)
  =
  \lim_{t\to\infty}\omega_0(\tau_t(A)),
  \label{eq:ness-local-weak-limit}
\end{equation}
if the limit exists.  The limit is local: it is not a statement that the
entire infinite-volume density matrix converges in trace norm.  When the
large-time state has a ray dependence, one instead studies limits of
\(\omega_0(\tau_t(A_{x=vt}))\), with \(v=x/t\).  The central steady state is
the ray \(v=0\).

The physical currents in the central steady state are defined by the
expectation values of local current operators:
\[
  J_E=\omega_{\rm ss}(j_E),
  \qquad
  J_Q=\omega_{\rm ss}(j_Q),
\]
with the convention that positive current flows from left to right.

\section{Reservoir Entropy Production}

Let the right reservoir gain energy and charge at rates
\[
  \frac{dE_R}{dt}=J_E,\qquad
  \frac{dQ_R}{dt}=J_Q,
\]
while the left reservoir loses the same amounts.  The reservoir entropy
change is computed using the equilibrium reservoir first laws
\[
  dS_\alpha=\beta_\alpha(dE_\alpha-\mu_\alpha dQ_\alpha),
  \qquad
  \alpha=L,R.
\]
Therefore
\begin{align}
  \dot S_{\rm res}
  &=
  \beta_R(J_E-\mu_RJ_Q)
  +\beta_L(-J_E+\mu_LJ_Q)
  \nonumber\\
  &=
  (\beta_R-\beta_L)J_E
  -(\beta_R\mu_R-\beta_L\mu_L)J_Q .
  \label{eq:ness-reservoir-entropy-production}
\end{align}
This formula is an identity once the currents and reservoir parameters are
defined.  Positivity is a dynamical statement about the state and the
reservoir construction.  In linear response, if
\[
  X=
  \begin{pmatrix}
    \beta_R-\beta_L\\
    -(\beta_R\mu_R-\beta_L\mu_L)
  \end{pmatrix},
  \qquad
  \begin{pmatrix}J_E\\ J_Q\end{pmatrix}
  =
  L X,
\]
then \(\dot S_{\rm res}=X^TLX\).  The second law in this linear regime is
equivalent to the positive semidefiniteness of the symmetric part of the
Onsager matrix \(L\).

\section{Finite Local Detailed Balance}
\label{sec:finite-local-detailed-balance-ness}

The reservoir formula above is the steady-state remnant of a stronger
finite statement.  We record it because it fixes the sign convention for
entropy production before any continuum or hydrodynamic limit is taken.

\begin{definition}[Finite reservoir jump datum]
\label{def:finite-reservoir-jump-datum}
A finite reservoir jump datum consists of:
\begin{enumerate}
\item a finite set \(I\) of system configurations;
\item real energy and charge functions \(E_i,Q_i\) on \(I\);
\item reservoirs \(\alpha\) with parameters \((\beta_\alpha,\mu_\alpha)\);
\item transition rates \(W^{(\alpha)}_{ij}\ge0\), \(i\ne j\), for jumps
  \(i\to j\) induced by reservoir \(\alpha\).
\end{enumerate}
The rates obey local detailed balance when
\begin{equation}
  W^{(\alpha)}_{ij}>0
  \quad\Longleftrightarrow\quad
  W^{(\alpha)}_{ji}>0,
  \qquad
  \log\frac{W^{(\alpha)}_{ij}}{W^{(\alpha)}_{ji}}
  =
  -\beta_\alpha
  \left[
    E_j-E_i-\mu_\alpha(Q_j-Q_i)
  \right].
  \label{eq:finite-local-detailed-balance}
\end{equation}
The master equation is
\begin{equation}
  \dot p_i
  =
  \sum_{\alpha}\sum_{j\ne i}
  \left[
    p_jW^{(\alpha)}_{ji}-p_iW^{(\alpha)}_{ij}
  \right],
  \qquad p_i\ge0,\qquad \sum_i p_i=1 .
  \label{eq:finite-reservoir-master-equation}
\end{equation}
\end{definition}

\paragraph{Finite entropy production.}
For a finite reservoir jump datum satisfying
\eqref{eq:finite-local-detailed-balance}, define the Shannon entropy
\begin{equation}
  S_{\rm sys}(p)=-\sum_i p_i\log p_i .
  \label{eq:finite-ness-system-entropy}
\end{equation}
The reservoir entropy-production rate is
\begin{equation}
  \dot S_{\rm res}
  =
  \sum_{\alpha}\sum_{i\ne j}
  p_iW^{(\alpha)}_{ij}
  \log\frac{W^{(\alpha)}_{ij}}{W^{(\alpha)}_{ji}},
  \label{eq:finite-ness-reservoir-entropy-rate}
\end{equation}
and the total entropy-production rate is
\begin{equation}
  \dot S_{\rm tot}
  =
  \frac12\sum_{\alpha}\sum_{i\ne j}
  \left(
    p_iW^{(\alpha)}_{ij}
    -
    p_jW^{(\alpha)}_{ji}
  \right)
  \log
  \frac{p_iW^{(\alpha)}_{ij}}
       {p_jW^{(\alpha)}_{ji}}
  \ge0 .
  \label{eq:finite-ness-total-entropy-production}
\end{equation}
Moreover
\begin{equation}
  \dot S_{\rm tot}=\frac{d}{dt}S_{\rm sys}(p)+\dot S_{\rm res}.
  \label{eq:finite-ness-system-plus-reservoir}
\end{equation}
At a stationary distribution \(p^{\rm ss}\), the system entropy derivative
vanishes, hence \(\dot S_{\rm res}=\dot S_{\rm tot}\ge0\).
The local detailed-balance relation says that a jump \(i\to j\) changes the
reservoir entropy by
\begin{equation}
  -\beta_\alpha
  \left[
    E_j-E_i-\mu_\alpha(Q_j-Q_i)
  \right]
  =
  \log\frac{W^{(\alpha)}_{ij}}{W^{(\alpha)}_{ji}},
  \label{eq:finite-ness-jump-reservoir-entropy}
\end{equation}
which gives \eqref{eq:finite-ness-reservoir-entropy-rate} after averaging
over jumps.  Differentiating \(S_{\rm sys}\) and using
\(\sum_i\dot p_i=0\) gives
\begin{equation}
  \frac{dS_{\rm sys}}{dt}
  =
  -\sum_i\dot p_i\log p_i .
  \label{eq:finite-ness-system-entropy-derivative}
\end{equation}
Substituting the master equation and symmetrizing over \(i,j\),
\begin{equation}
  \frac{dS_{\rm sys}}{dt}
  =
  \frac12\sum_{\alpha}\sum_{i\ne j}
  \left(
    p_iW^{(\alpha)}_{ij}
    -
    p_jW^{(\alpha)}_{ji}
  \right)
  \log\frac{p_i}{p_j}.
  \label{eq:finite-ness-system-entropy-current-form}
\end{equation}
Similarly, \eqref{eq:finite-ness-reservoir-entropy-rate} can be written as
\begin{equation}
  \dot S_{\rm res}
  =
  \frac12\sum_{\alpha}\sum_{i\ne j}
  \left(
    p_iW^{(\alpha)}_{ij}
    -
    p_jW^{(\alpha)}_{ji}
  \right)
  \log\frac{W^{(\alpha)}_{ij}}{W^{(\alpha)}_{ji}} .
  \label{eq:finite-ness-reservoir-current-form}
\end{equation}
Adding the last two equations gives
\eqref{eq:finite-ness-total-entropy-production}.  Each summand is
nonnegative because
\begin{equation}
  (x-y)\log(x/y)\ge0,\qquad x>0,\quad y>0 .
  \label{eq:finite-ness-log-monotonicity}
\end{equation}
Terms with \(x=y=0\) are omitted by continuity.  This proves both
\eqref{eq:finite-ness-system-plus-reservoir} and the inequality.

The two-reservoir current formula
\eqref{eq:ness-reservoir-entropy-production} is obtained from this finite
statement when the stationary currents have well-defined macroscopic energy
and charge rates.  The finite theorem does not prove existence of the
long-time QFT state \(\omega_{\rm ss}\); it proves the algebraic positivity
that any Markovian reservoir reduction with local detailed balance must
satisfy.

\section{Finite Path Measures and Fluctuation Relations}
\label{sec:finite-path-measure-fluctuation-relations}

Fluctuation relations are identities between probability measures on
histories.  They become mathematical statements after the path space and the
reversed path measure have been constructed.  We state the finite
continuous-time version because it is the regulator-level statement that
survives, if the corresponding continuum limit exists, in nonequilibrium QFT.

Let \(I\) be a finite configuration set.  For \(0\leq t\leq T\), let
\(W^{(\alpha)}_{ij}(t)\geq0\), \(i\ne j\), be time-dependent jump rates
labelled by reservoirs \(\alpha\), and let
\[
  r_i(t)=\sum_{\alpha}\sum_{j\ne i}W^{(\alpha)}_{ij}(t)
\]
be the escape rate from \(i\).  A labelled trajectory is
\[
  \gamma=(i_0;t_1,\alpha_1,i_1;\ldots;t_n,\alpha_n,i_n),
  \qquad 0<t_1<\cdots<t_n<T,
\]
where the state is \(i_m\) on \(t_m<t<t_{m+1}\), with
\(t_0=0\) and \(t_{n+1}=T\).  Given an initial distribution \(p_0\), the
forward path density with respect to counting measure over labels and
Lebesgue measure over ordered jump times is
\begin{align}
  \mathbb P_F(\gamma)
  &=
  p_0(i_0)
  \exp\!\left[-\int_{0}^{t_1}r_{i_0}(s)\,ds\right]
  \notag\\
  &\quad\times
  \prod_{m=1}^{n}
  \left[
    W^{(\alpha_m)}_{i_{m-1}i_m}(t_m)
    \exp\!\left[-\int_{t_m}^{t_{m+1}}r_{i_m}(s)\,ds\right]
  \right].
  \label{eq:finite-forward-jump-path-density}
\end{align}
The reversed protocol has rates
\[
  \widetilde W^{(\alpha)}_{ij}(t)=W^{(\alpha)}_{ij}(T-t)
\]
and initial distribution \(q_0\).  The time-reversed trajectory is
\[
  \Theta\gamma
  =
  (i_n;T-t_n,\alpha_n,i_{n-1};\ldots;T-t_1,\alpha_1,i_0).
\]
Assume absolute continuity: whenever a forward path has positive
probability, the reversed path has positive probability under the reversed
protocol.  Define the trajectory entropy production by the
Radon--Nikodym logarithm
\begin{equation}
  \Sigma(\gamma)
  =
  \log
  \frac{\mathbb P_F(\gamma)}
       {\mathbb P_R(\Theta\gamma)} .
  \label{eq:finite-path-entropy-production-rn}
\end{equation}

\begin{proposition}[Finite path-measure fluctuation identity]
\label{prop:finite-path-fluctuation-identity}
For the finite jump process above,
\begin{align}
  \Sigma(\gamma)
  &=
  \log\frac{p_0(i_0)}{q_0(i_n)}
  +
  \sum_{m=1}^n
  \log
  \frac{W^{(\alpha_m)}_{i_{m-1}i_m}(t_m)}
       {W^{(\alpha_m)}_{i_m i_{m-1}}(t_m)},
  \label{eq:finite-path-entropy-explicit}\\
  \left\langle e^{-\Sigma}\right\rangle_F&=1 .
  \label{eq:finite-integral-fluctuation-theorem}
\end{align}
If \(q_0\) is chosen to be the forward final distribution \(p_T\), then
\(\langle \Sigma\rangle_F\ge0\).  More generally, \(\langle\Sigma\rangle_F\)
is the relative entropy
\[
  D(\mathbb P_F\Vert \Theta^{-1}\mathbb P_R)
\]
between the forward history measure and the pullback of the reversed history
measure.
\end{proposition}

\begin{proof}
Insert the path densities.  The waiting-time factors in the denominator are
the same factors as in the numerator after the change of variables
\(s\mapsto T-s\) on each reversed interval.  They therefore cancel.  What
remains is exactly the initial-distribution ratio and the product of jump
rate ratios in \eqref{eq:finite-path-entropy-explicit}.

The integral identity is the normalization of the reversed path measure:
\[
  \left\langle e^{-\Sigma}\right\rangle_F
  =
  \int e^{-\Sigma(\gamma)}\,d\mathbb P_F(\gamma)
  =
  \int d\mathbb P_R(\Theta\gamma)
  =
  1 .
\]
The map \(\Theta\) is an involution on the labelled path space, so the last
integral is the total mass of \(\mathbb P_R\).  Finally,
\[
  \langle\Sigma\rangle_F
  =
  \int
  \log\frac{d\mathbb P_F}{d(\Theta^{-1}\mathbb P_R)}
  \,d\mathbb P_F
  =
  D(\mathbb P_F\Vert\Theta^{-1}\mathbb P_R)\ge0,
\]
with equality precisely when the two history measures agree almost surely.
When \(q_0=p_T\), the boundary term in
\eqref{eq:finite-path-entropy-explicit} is the stochastic system entropy
change \(-\log p_T(i_n)+\log p_0(i_0)\), and the jump-rate term is the
reservoir entropy change under local detailed balance.  Hence the average
is the mean total entropy production.
\end{proof}

The detailed fluctuation relation is a direct disintegration of the same
identity.  If \(P_F\) is the forward law of \(\Sigma\) and \(P_R\) is the
reversed law of the reversed entropy production, then
\[
  P_F(ds)=e^{s}\,P_R(-ds)
\]
on every Borel set on which both measures are defined.  This form is often
the most convenient one experimentally, but the Radon--Nikodym statement
\eqref{eq:finite-path-entropy-production-rn} is the mathematical content.

\paragraph{Jarzynski identity as a special case.}
Take one thermal reservoir with inverse temperature \(\beta\), no chemical
potential, and a time-dependent finite energy function \(E_i(t)\).  Assume
local detailed balance at each jump time:
\[
  \log\frac{W_{ij}(t)}{W_{ji}(t)}
  =
  -\beta\bigl(E_j(t)-E_i(t)\bigr).
\]
Let \(p_0(i)=Z(0)^{-1}e^{-\beta E_i(0)}\) and
\(q_0(i)=Z(T)^{-1}e^{-\beta E_i(T)}\).  For a trajectory \(\gamma\), define
the protocol work
\begin{equation}
  W_{\rm ext}(\gamma)
  =
  \int_0^T \partial_tE_{i(t)}(t)\,dt,
  \label{eq:finite-path-protocol-work}
\end{equation}
where \(i(t)\) is the piecewise-constant state path.  The total system
energy change decomposes as
\[
  E_{i_n}(T)-E_{i_0}(0)
  =
  W_{\rm ext}(\gamma)
  +
  \sum_{m=1}^n
  \bigl(E_{i_m}(t_m)-E_{i_{m-1}}(t_m)\bigr).
\]
Combining this identity with local detailed balance gives
\[
  \Sigma(\gamma)
  =
  \beta\bigl(W_{\rm ext}(\gamma)-\Delta F\bigr),
  \qquad
  \Delta F=F(T)-F(0),
  \qquad
  F(t)=-\beta^{-1}\log Z(t).
\]
Therefore \eqref{eq:finite-integral-fluctuation-theorem} becomes
\begin{equation}
  \left\langle e^{-\beta W_{\rm ext}}\right\rangle_F
  =
  e^{-\beta\Delta F}.
  \label{eq:finite-path-jarzynski}
\end{equation}
For a continuum QFT quench, the corresponding use of
\eqref{eq:finite-path-jarzynski} requires a specified finite regulator, a
measured work protocol or two-time energy measurement replacement, a
time-reversal operation, and a limit in which the history measures converge.

\section{Finite-Step Langevin Measures and MSRJD Variables}
\label{sec:finite-step-langevin-msrjd}

The Martin--Siggia--Rose--Janssen--de Dominicis (MSRJD) response-field path
integral used in classical stochastic field theory is an exact rewriting of
a finite-dimensional path measure before it is a continuum notation.  Let
\(x_n=(x_n^1,\ldots,x_n^N)\in\mathbb R^N\) be the variables of a finite
regulator at \(t_n=n\Delta t\), \(0\le n\le M\).  The index
\(a=1,\ldots,N\) may include internal labels and lattice sites.  Fix a smooth
drift vector field \(F^a(x,t)\) and a real symmetric positive definite matrix
\(D^{ab}\).  The Ito finite-step process is
\begin{equation}
  x_{n+1}^a
  =
  x_n^a+\Delta t\,F^a(x_n,t_n)+\eta_n^a,
  \qquad
  \mathbb E(\eta_n^a\eta_m^b)
  =
  2\Delta t\,D^{ab}\delta_{nm}.
  \label{eq:finite-langevin-ito-step}
\end{equation}
With \(x_0\) fixed and
\[
  \Delta x_n^a=x_{n+1}^a-x_n^a,
  \qquad
  F_n^a=F^a(x_n,t_n),
\]
the conditional path density is
\begin{equation}
  \mathbb P_\Delta[x_1,\ldots,x_M\mid x_0]
  =
  \prod_{n=0}^{M-1}
  \frac{
    \exp\!\left[
      -\frac{1}{4\Delta t}
      (\Delta x_n-\Delta t F_n)_a(D^{-1})^{ab}
      (\Delta x_n-\Delta t F_n)_b
    \right]}
    {\sqrt{(4\pi\Delta t)^N\det D}} .
  \label{eq:finite-langevin-path-density}
\end{equation}
This formula is the regulated stochastic theory.  A continuum stochastic
field theory is a limit of such finite measures, together with convergence
of the observables under study.

The Gaussian kernel in \eqref{eq:finite-langevin-path-density} has the
Fourier representation
\begin{equation}
  \frac{
    \exp\!\left[-\frac{1}{4\Delta t}y_a(D^{-1})^{ab}y_b\right]}
    {\sqrt{(4\pi\Delta t)^N\det D}}
  =
  \int_{\mathbb R^N}\frac{d^N\hat x}{(2\pi)^N}
  \exp\!\left[
    -\ii\hat x_a y^a
    -\Delta t\,\hat x_aD^{ab}\hat x_b
  \right].
  \label{eq:finite-msrjd-gaussian-fourier}
\end{equation}
Indeed, the right side is the elementary Gaussian integral
\[
  \frac{1}{(2\pi)^N}
  \int_{\mathbb R^N}d^N\hat x\,
  \exp\!\left[-\hat x_a(\Delta t D^{ab})\hat x_b-\ii\hat x_a y^a\right],
\]
and completing the square gives the determinant and exponent on the left.
The finite-step MSRJD identity follows by applying
\eqref{eq:finite-msrjd-gaussian-fourier} to
\(y^a=\Delta x_n^a-\Delta t F_n^a\) at each time step:
\begin{align}
  \mathbb E_\Delta[\mathcal O]
  &=
  \int_{\mathbb R^{NM}}\prod_{n=1}^{M}d^Nx_n\,
  \mathcal O(x_0,\ldots,x_M)
  \prod_{n=0}^{M-1}
  \int_{\mathbb R^N}\frac{d^N\hat x_n}{(2\pi)^N}
  \notag\\
  &\qquad\qquad\times
  \exp\!\left\{
    \sum_{n=0}^{M-1}
    \left[
      -\ii\hat x_{n,a}
      \bigl(\Delta x_n^a-\Delta t F_n^a\bigr)
      -\Delta t\,\hat x_{n,a}D^{ab}\hat x_{n,b}
    \right]
  \right\}.
  \label{eq:finite-msrjd-identity}
\end{align}
Here \(\mathcal O\) is any bounded measurable functional of the finite path.
If the initial condition has density \(\rho_0(x_0)\), one also integrates
over \(x_0\) with weight \(\rho_0\).  The auxiliary variable
\(\hat x_{n,a}\) is the Fourier-dual variable to the finite increment
constraint.  The frequently used variable \(\widetilde x_{n,a}=\ii\hat x_{n,a}\)
is a contour notation for the same finite integral; its convergence and
allowed deformations are inherited from \eqref{eq:finite-msrjd-gaussian-fourier}.

The same finite construction fixes the generator and the Fokker--Planck
equation from finite moments.  For
\(f\in C_b^4(\mathbb R^N)\), Taylor expansion of
\(f(x+\Delta t F(x,t)+\eta)\) and the Gaussian moments
\(\mathbb E\eta^a=0\),
\(\mathbb E(\eta^a\eta^b)=2\Delta t D^{ab}\) give
\begin{equation}
  \mathbb E[f(x_{n+1})\mid x_n=x]
  =
  f(x)
  +
  \Delta t
  \left(F^a(x,t_n)\partial_a f(x)
  +D^{ab}\partial_a\partial_b f(x)\right)
  +O((\Delta t)^2)
  \label{eq:finite-langevin-generator-expansion}
\end{equation}
uniformly on compact sets where the drift and the required derivatives of
\(f\) are bounded.  The limiting Markov generator on test functions is
therefore
\begin{equation}
  L_t f=F^a(x,t)\partial_a f+D^{ab}\partial_a\partial_b f .
  \label{eq:finite-langevin-generator}
\end{equation}
The adjoint equation for a smooth density \(p(x,t)\) is
\begin{equation}
  \partial_t p
  =
  -\partial_a(F^a p)
  +D^{ab}\partial_a\partial_b p .
  \label{eq:finite-langevin-fokker-planck}
\end{equation}
For stochastic PDEs and QFT applications, \(\mathbb R^N\) is replaced first
by a finite spatial regulator and only then by a continuum limit.  The data
to be controlled are the regulated drift, the regulated covariance, the
choice of stochastic discretization, ultraviolet renormalizations of drift
and noise, and convergence of correlation functions or path measures.

\section{Finite Reaction Networks and Doi--Peliti Variables}
\label{sec:finite-reaction-network-doi-peliti}

Doi--Peliti variables are the occupation-number analogue of the response
variables above.  The starting point is a Markov generator on a countable
occupation basis.  Let \(s=1,\ldots,S\) label a finite set of species, and
let \(n=(n_1,\ldots,n_S)\in\mathbb N^S\) be the occupation vector of a finite
spatial regulator.  For a multi-index \(\alpha\in\mathbb N^S\), set
\[
  z^\alpha=\prod_s z_s^{\alpha_s},
  \qquad
  (n)_\alpha=\prod_s n_s(n_s-1)\cdots(n_s-\alpha_s+1),
\]
with \((n)_\alpha=0\) if some \(n_s<\alpha_s\).  A reaction \(r\) has input
\(\nu_r\), output \(\nu'_r\), stoichiometric vector
\(v_r=\nu'_r-\nu_r\), and rate constant \(\kappa_r\ge0\).  The mass-action
jump rate is
\[
  w_r(n)=\kappa_r(n)_{\nu_r},
  \qquad
  n\longmapsto n+v_r .
\]
The master equation for the probability \(P_t(n)\) is
\begin{equation}
  \partial_tP_t(m)
  =
  \sum_r
  \left[
    \kappa_r(m-v_r)_{\nu_r}P_t(m-v_r)
    -
    \kappa_r(m)_{\nu_r}P_t(m)
  \right],
  \label{eq:reaction-network-master-equation}
\end{equation}
where terms with \(m-v_r\notin\mathbb N^S\) vanish.

Let \(\mathcal F_{\rm alg}\) be the algebraic bosonic Fock space with basis
\(\{|n\rangle:n\in\mathbb N^S\}\).  Define
\[
  a_s|n\rangle=n_s|n-e_s\rangle,
  \qquad
  a_s^\dagger|n\rangle=|n+e_s\rangle,
  \qquad
  [a_s,a_t^\dagger]=\delta_{st}.
\]
Here \(a_s|n\rangle=0\) when \(n_s=0\).
For \(\alpha\in\mathbb N^S\), write
\[
  a^\alpha=\prod_s a_s^{\alpha_s},
  \qquad
  (a^\dagger)^\alpha=\prod_s(a_s^\dagger)^{\alpha_s}.
\]
Encode a probability distribution as
\[
  |P_t\rangle=\sum_{n\in\mathbb N^S}P_t(n)|n\rangle .
\]
The Doi--Peliti generator is the normally ordered operator
\begin{equation}
  \mathcal L_{\rm DP}
  =
  \sum_r
  \kappa_r
  \left[
    (a^\dagger)^{\nu'_r}
    -
    (a^\dagger)^{\nu_r}
  \right]
  a^{\nu_r}.
  \label{eq:doi-peliti-generator}
\end{equation}

\begin{proposition}[Finite Doi--Peliti generator identity]
\label{prop:finite-doi-peliti-generator-identity}
For every finitely supported probability vector \(P\), the coefficient of
\(|m\rangle\) in \(\mathcal L_{\rm DP}|P\rangle\) is the right side of
\eqref{eq:reaction-network-master-equation}.  Therefore
\(\partial_t|P_t\rangle=\mathcal L_{\rm DP}|P_t\rangle\) is exactly the
reaction-network master equation on the algebraic domain.  If
\(\langle{\bf 1}|n\rangle=1\) for every \(n\), then
\[
  \langle{\bf 1}|\mathcal L_{\rm DP}=0 .
\]
\end{proposition}

\begin{proof}
For every multi-index \(\alpha\),
\[
  a^\alpha|n\rangle=(n)_\alpha|n-\alpha\rangle.
\]
Hence a single reaction term acts by
\[
  \kappa_r
  \left[
    (a^\dagger)^{\nu'_r}
    -
    (a^\dagger)^{\nu_r}
  \right]
  a^{\nu_r}|n\rangle
  =
  \kappa_r(n)_{\nu_r}
  \left(|n+v_r\rangle-|n\rangle\right).
\]
After multiplying by \(P(n)\) and summing over \(n\), the coefficient of
\(|m\rangle\) is the gain from \(n=m-v_r\) minus the loss from \(n=m\),
which is \eqref{eq:reaction-network-master-equation}.  Finally,
\(\langle{\bf 1}|(a^\dagger)^\alpha=\langle{\bf 1}|\) for every
\(\alpha\), because both sides assign value \(1\) to every basis vector.
Thus each bracket in \eqref{eq:doi-peliti-generator} is annihilated by
\(\langle{\bf 1}|\), giving probability conservation.
\end{proof}

The coherent-state form is a controlled time-slicing of the same operator.
Let
\[
  |z\rangle=\exp\!\left(\sum_s z_sa_s^\dagger\right)|0\rangle,
  \qquad
  \langle\bar z|=\langle0|\exp\!\left(\sum_s \bar z_sa_s\right),
\]
where \(z,\bar z\in\mathbb C^S\).  Then
\[
  a_s|z\rangle=z_s|z\rangle,
  \qquad
  \langle\bar z|a_s^\dagger=\bar z_s\langle\bar z|,
  \qquad
  \langle\bar z|z\rangle=e^{\bar z\cdot z}.
\]
The normal symbol of the generator is
\begin{equation}
  H_{\rm DP}(\bar z,z)
  =
  e^{-\bar z\cdot z}
  \langle\bar z|\mathcal L_{\rm DP}|z\rangle
  =
  \sum_r
  \kappa_r
  \left(\bar z^{\nu'_r}-\bar z^{\nu_r}\right)z^{\nu_r}.
  \label{eq:doi-peliti-normal-symbol}
\end{equation}
In any finite-dimensional truncation whose projected generator and boundary
rules have been specified, and in particular in any finite total-number
sector preserved by the reactions, the coherent-state resolution of the
identity gives an ordinary finite-dimensional time-sliced integral for
matrix elements of \(\exp(T\mathcal L_{\rm DP})\).  Replacing each
short-time factor by
\[
  \langle\bar z_{j+1}|e^{\Delta t\mathcal L_{\rm DP}}|z_j\rangle
  =
  \exp(\bar z_{j+1}\cdot z_j)
  \left[
    1+\Delta t\,H_{\rm DP}(\bar z_{j+1},z_j)+O((\Delta t)^2)
  \right]
\]
gives the continuum Doi--Peliti action whenever the time-slicing limit is
controlled.  The shift \(\bar z=1+\widetilde z\) is useful because
\[
  H_{\rm DP}(1,z)=0,
\]
which is the symbol form of probability conservation.  Differentiating at
\(\bar z=1\) gives the deterministic rate equation for the mean-field
density variables:
\begin{equation}
  \dot z_s
  =
  \left.
  \frac{\partial H_{\rm DP}}{\partial\bar z_s}
  \right|_{\bar z=1}
  =
  \sum_r
  \kappa_r(\nu'_{r,s}-\nu_{r,s})z^{\nu_r}.
  \label{eq:doi-peliti-mean-field-rate-equation}
\end{equation}

The same finite generator also exposes the dynamical large-deviation
Hamiltonian.  For an exponential test function
\[
  f_p(n)=e^{p\cdot n},
\]
the backward generator \(L\) associated with
\eqref{eq:reaction-network-master-equation} satisfies
\begin{equation}
  \frac{(Lf_p)(n)}{f_p(n)}
  =
  \sum_r
  \kappa_r(n)_{\nu_r}
  \left(e^{p\cdot v_r}-1\right).
  \label{eq:reaction-network-exponential-test-function}
\end{equation}
This follows by applying the backward generator to \(f_p\):
the \(r\)-th jump changes \(f_p(n)\) to
\(f_p(n+v_r)=e^{p\cdot v_r}f_p(n)\), while the loss term subtracts
\(f_p(n)\).
If a density scaling \(n=\Omega x\) is chosen together with rate constants
scaled so that \(\kappa_r(n)_{\nu_r}=\Omega\,\lambda_r(x)+o(\Omega)\), then
\[
  \mathcal H_{\rm LD}(x,p)
  =
  \sum_r
  \lambda_r(x)
  \left(e^{p\cdot v_r}-1\right)
\]
is the finite-regulator Hamiltonian for the path rate function.  Proving the
large-deviation principle itself requires tightness, exponential estimates,
and a specified scaling limit;
\eqref{eq:reaction-network-exponential-test-function} is the exact algebraic
input.

\section{Tilted Generators and Dynamical Large Deviations}
\label{sec:finite-tilted-generators-large-deviations}

Finite path measures also give the precise source of dynamical
large-deviation generators.  Let \(I\) be a finite state space and let
\(W_{ij}\ge0\), \(i\ne j\), be time-independent jump rates, with
\[
  r_i=\sum_{j\ne i}W_{ij}.
\]
For a trajectory \(i(t)\), define an additive observable
\begin{equation}
  A_T(\gamma)
  =
  \int_0^T h_{i(t)}\,dt
  +
  \sum_{m=1}^{N_T} g_{i_{m-1}i_m},
  \label{eq:finite-additive-jump-observable}
\end{equation}
where \(h_i\in\mathbb R\) is a state weight, \(g_{ij}\in\mathbb R\) is a
jump weight, and \(N_T\) is the number of jumps up to time \(T\).  For
\(\lambda\in\mathbb C\), the tilted backward generator on functions
\(f:I\to\mathbb C\) is
\begin{equation}
  (\mathsf L_\lambda f)_i
  =
  \lambda h_i f_i
  +
  \sum_{j\ne i}
  W_{ij}
  \left(e^{\lambda g_{ij}}f_j-f_i\right).
  \label{eq:finite-tilted-generator}
\end{equation}

\begin{proposition}[Finite Feynman--Kac identity]
\label{prop:finite-feynman-kac-tilted-generator}
For \(i\in I\), set
\[
  u_i(t)=\mathbb E_i[e^{\lambda A_t}f(i(t))],
\]
where the chain starts from \(i\) at time \(0\).  Then
\[
  \partial_tu(t)=\mathsf L_\lambda u(t),
  \qquad
  u(0)=f .
\]
Consequently, for an initial distribution \(p_0\),
\begin{equation}
  \mathbb E_{p_0}\!\left[e^{\lambda A_T}\right]
  =
  \sum_{i\in I}p_0(i)\,
  \bigl(e^{T\mathsf L_\lambda}{\bf 1}\bigr)_i .
  \label{eq:finite-feynman-kac-mgf}
\end{equation}
\end{proposition}

\begin{proof}
Condition on the evolution during a short interval \([0,\Delta t]\).  With
probability \(1-r_i\Delta t+O((\Delta t)^2)\), the chain stays at \(i\), and
the additive functional contributes \(h_i\Delta t+O((\Delta t)^2)\).  With
probability \(W_{ij}\Delta t+O((\Delta t)^2)\), the chain jumps once from
\(i\) to \(j\), contributing the jump weight \(g_{ij}\) and the state weight
over a time interval of length \(\Delta t\).  In the jump branch the latter
changes the contribution only by \(1+O(\Delta t)\), and hence only by
\(O((\Delta t)^2)\) after multiplication by the jump probability.  Therefore
\[
  u_i(t+\Delta t)
  =
  \left(1+\lambda h_i\Delta t-r_i\Delta t\right)u_i(t)
  +
  \sum_{j\ne i}
  W_{ij}\Delta t\,e^{\lambda g_{ij}}u_j(t)
  +
  O((\Delta t)^2),
\]
uniformly on the finite state space.  Dividing by \(\Delta t\) and sending
\(\Delta t\to0\) gives \(\partial_tu=\mathsf L_\lambda u\).  The final
identity follows by taking \(f={\bf 1}\) and averaging over \(p_0\).
\end{proof}

If the original Markov chain is irreducible, then
\(\mathsf L_\lambda+c{\bf 1}\) is an irreducible nonnegative matrix for
real \(\lambda\) and all sufficiently large \(c\).  Perron--Frobenius
therefore gives a simple real eigenvalue with strictly positive left and
right eigenvectors and with maximal real part.  It identifies the long-time
scaled cumulant generating function
\[
  \psi(\lambda)
  =
  \lim_{T\to\infty}
  \frac1T\log
  \mathbb E_{p_0}\!\left[e^{\lambda A_T}\right]
\]
with the eigenvalue of \(\mathsf L_\lambda\) whose real part is maximal,
independently of any nonzero nonnegative initial distribution.  The
positivity of the Perron eigenvectors is what prevents the leading spectral
coefficient in \eqref{eq:finite-feynman-kac-mgf} from vanishing.  A path
large-deviation principle and a Legendre-dual rate function require the
separate analytic and tightness hypotheses of large-deviation theory.  The
finite tilted generator is the exact microscopic input to those hypotheses.

There is a useful exact symmetry when \(A_T\) is entropy production in a
stationary finite jump process.  Assume \(W_{ij}>0\) if and only if
\(W_{ji}>0\), and let \(\pi_i>0\) be the stationary distribution.  Define
the entropy increment of one jump by
\begin{equation}
  \sigma_{ij}
  =
  \log\frac{\pi_iW_{ij}}{\pi_jW_{ji}} .
  \label{eq:stationary-entropy-jump-increment}
\end{equation}
Let \(\mathsf L_q^\Sigma\) be the tilted generator for
\(\mathbb E[e^{-q\Sigma_T}]\):
\begin{equation}
  (\mathsf L_q^\Sigma f)_i
  =
  \sum_{j\ne i}
  W_{ij}e^{-q\sigma_{ij}}f_j-r_if_i .
  \label{eq:entropy-production-tilted-generator}
\end{equation}
With \(\Pi=\operatorname{diag}(\pi_i)\), one has the finite matrix identity
\begin{equation}
  \mathsf L_q^\Sigma
  =
  \Pi^{-1}
  \bigl(\mathsf L_{1-q}^\Sigma\bigr)^T
  \Pi .
  \label{eq:finite-gallavotti-cohen-similarity}
\end{equation}
Indeed, the off-diagonal \(i,j\) entry of the right side is
\[
  \frac{\pi_j}{\pi_i}
  W_{ji}e^{-(1-q)\sigma_{ji}}
  =
  W_{ij}
  \left(\frac{\pi_jW_{ji}}{\pi_iW_{ij}}\right)^q
  =
  W_{ij}e^{-q\sigma_{ij}},
\]
and the diagonal entries are both \(-r_i\).  Hence
\(\mathsf L_q^\Sigma\) and \(\mathsf L_{1-q}^\Sigma\) have the same
spectrum, and the entropy-production scaled cumulant generating function
obeys
\[
  \psi_\Sigma(q)=\psi_\Sigma(1-q).
\]
This is the finite-generator form of the Gallavotti--Cohen symmetry.  The
finite-time fluctuation relation in
Section~\ref{sec:finite-path-measure-fluctuation-relations} is the
corresponding path-measure identity; the spectral symmetry is its
time-homogeneous stationary long-time shadow.

\section{Weak-Coupling Open-System Limit}

Let a system Hilbert space couple to an environment by
\begin{equation}
  H=H_S+H_E+\lambda\sum_a S_a\otimes E_a .
  \label{eq:open-system-total-hamiltonian}
\end{equation}
We take \(S_a=S_a^\dagger\) and \(E_a=E_a^\dagger\); a non-Hermitian
coupling is rewritten in this form by separating Hermitian and
anti-Hermitian parts.  This convention fixes the adjoints in the spectral
matrices below.
Assume the environment state \(\rho_E\) is stationary,
\([H_E,\rho_E]=0\), and that connected environment correlators decay
fast enough for their Fourier transforms to exist as positive matrices:
\begin{equation}
  \gamma_{ab}(\omega)
  =
  \int_{-\infty}^{\infty}dt\,
  e^{\ii\omega t}
  \operatorname{Tr}_E\!\left(
  \rho_E E_a(t)E_b(0)
  \right).
  \label{eq:open-system-bath-spectral-matrix}
\end{equation}
Positivity follows because, for any complex vector \(v_a\),
\[
  C_v(t)=
  v_a^*v_b\,\operatorname{Tr}_E\!\left(\rho_EE_a(t)E_b(0)\right)
  =
  \operatorname{Tr}_E\!\left(\rho_EF(t)^\dagger F(0)\right),
  \qquad F=v_bE_b ,
\]
is a function of positive type.  Indeed, for any times \(t_i\) and complex
numbers \(c_i\),
\[
  \sum_{i,j}c_i^*c_j\,C_v(t_i-t_j)
  =
  \operatorname{Tr}_E\!\left(
    \rho_E
    \left(\sum_i c_iF(t_i)\right)^\dagger
    \left(\sum_j c_jF(t_j)\right)
  \right)\ge0.
\]
The distributional Bochner theorem then says that its Fourier transform is a
positive measure.  Hence \(v_a^*\gamma_{ab}(\omega)v_b\ge0\) at frequencies
where the density \(\gamma(\omega)\) exists, and more generally as a positive
matrix-valued measure.

Let \(P_\epsilon\) be the spectral projections of \(H_S\), and define
Bohr-frequency components
\begin{equation}
  S_a(\omega)
  =
  \sum_{\epsilon'-\epsilon=\omega}
  P_\epsilon S_aP_{\epsilon'} .
  \label{eq:open-system-bohr-components}
\end{equation}
Under the weak-coupling scaling
\[
  \lambda\to0,
  \qquad
  t=\lambda^{-2}\tau,
\]
the interaction-picture density matrix satisfies the integral equation
\[
  \rho_I(t)
  =
  \rho_I(0)
  -\ii\lambda\int_0^t ds\,[V_I(s),\rho_I(s)],
  \qquad
  V_I(s)=\sum_aS_a(s)\otimes E_a(s).
\]
Expanding once more, tracing over the environment, assuming
\(\operatorname{Tr}_E(\rho_EE_a)=0\) after absorbing one-point terms into
\(H_S\), and using fast decay of connected environment correlators gives a
second-order memory integral.  The van Hove limit keeps
\(\lambda^2t=\tau\) fixed.  Oscillatory factors
\(e^{\ii(\omega-\omega')s}\) then project onto equal Bohr frequencies,
while the time integral of the bath correlator supplies
\(\gamma_{ab}(\omega)\) and its principal-value Hilbert transform.  For a
finite-dimensional system with discrete Bohr spectrum, bounded system
couplings, a stationary environment state, and absolutely integrable
environment connected correlators with the mixing estimates needed to
discard higher cumulants in the van Hove limit, this limit gives
\begin{equation}
  \frac{d\rho_S}{d\tau}
  =
  \mathcal L(\rho_S),
  \label{eq:open-system-markov-master-equation}
\end{equation}
where
\begin{align}
  \mathcal L(\rho)
  &=
  -\ii[H_S+H_{\rm LS},\rho]
  \nonumber\\
  &\quad+
  \sum_{\omega}\sum_{a,b}\gamma_{ab}(\omega)
  \left[
    S_b(\omega)\rho S_a(\omega)^\dagger
    -\frac12\{S_a(\omega)^\dagger S_b(\omega),\rho\}
  \right].
  \label{eq:open-system-davies-generator}
\end{align}
\(H_{\rm LS}\) is the Lamb-shift Hamiltonian from the principal-value part
of the bath correlator.  The dissipative part has the GKSL form because each
matrix \(\gamma_{ab}(\omega)\) is positive semidefinite and hence can be
diagonalized into jump operators.

\begin{proposition}[Trace preservation and complete positivity]
\label{prop:gksl-trace-cp}
For any finite-dimensional system, the generator
\eqref{eq:open-system-davies-generator} preserves
\(\operatorname{Tr}\rho\), preserves Hermiticity, and generates a completely
positive trace-preserving semigroup.
\end{proposition}

\begin{proof}
Trace preservation follows term by term:
\[
  \operatorname{Tr}(L\rho L^\dagger)
  -\frac12\operatorname{Tr}(L^\dagger L\rho)
  -\frac12\operatorname{Tr}(\rho L^\dagger L)=0
\]
by cyclicity of the trace.  The commutator has zero trace and preserves
Hermiticity.  Since \(\gamma(\omega)\) is positive, write
\(\gamma_{ab}(\omega)=\sum_\alpha v_{\alpha a}(\omega)^*
v_{\alpha b}(\omega)\) and define
\[
  L_{\alpha\omega}=\sum_b v_{\alpha b}(\omega)S_b(\omega).
\]
Then the dissipator is a sum of standard Lindblad terms
\[
  L_{\alpha\omega}\rho L_{\alpha\omega}^\dagger
  -\frac12\{L_{\alpha\omega}^\dagger L_{\alpha\omega},\rho\}.
\]
Set \(H=H_S+H_{\rm LS}\),
\[
  \Gamma=\sum_{\alpha,\omega}
  L_{\alpha\omega}^\dagger L_{\alpha\omega},
  \qquad
  H_{\rm eff}=H-\frac{\ii}{2}\Gamma ,
\]
and define the completely positive maps
\[
  \mathcal K_t(\rho)=
  e^{-\ii H_{\rm eff}t}\rho e^{\ii H_{\rm eff}^\dagger t},
  \qquad
  \mathcal J(\rho)=
  \sum_{\alpha,\omega}
  L_{\alpha\omega}\rho L_{\alpha\omega}^\dagger .
\]
In finite dimension the Dyson--Phillips series
\begin{align}
  T_t
  &=
  \mathcal K_t
  +
  \sum_{n\ge1}
  \int_{0<s_1<\cdots<s_n<t}
  \mathcal K_{t-s_n}\mathcal J
  \mathcal K_{s_n-s_{n-1}}\cdots
  \mathcal J\mathcal K_{s_1}\,
  ds_1\cdots ds_n
  \label{eq:open-system-gksl-dyson}
\end{align}
converges in operator norm on the trace-class matrices.  Each summand is
completely positive, so \(T_t\) is completely positive.  Differentiating the
series gives \(dT_t/dt=\mathcal L T_t=T_t\mathcal L\) and \(T_0=\id\).
Since the trace of \(\mathcal L(\rho)\) vanishes for every \(\rho\),
\[
  \frac{d}{dt}\operatorname{Tr}T_t(\rho)
  =
  \operatorname{Tr}\mathcal L(T_t(\rho))=0,
\]
and therefore \(T_t\) is trace preserving.  This proves complete positivity
and trace preservation for the semigroup generated by
\eqref{eq:open-system-davies-generator}.
\end{proof}

For comparison with the infinitesimal Kraus form, a short time step
\(d\tau\) has
\[
  K_0=\mathbf 1-d\tau\left(\ii H
  +\frac12\Gamma\right),
  \qquad
  K_{\alpha\omega}=\sqrt{d\tau}\,L_{\alpha\omega}.
\]
This is a first-order local representation of the same generator; the
identity
\[
  K_0^\dagger K_0
  +
  \sum_{\alpha,\omega}
  K_{\alpha\omega}^\dagger K_{\alpha\omega}
  =
  \mathbf 1+O(d\tau^2)
\]
is the infinitesimal trace-preservation statement, while the norm-convergent
series \eqref{eq:open-system-gksl-dyson} is the finite-time object.

\section{Thermal Bath and Detailed Balance}

If the environment is a KMS state at inverse temperature \(\beta\), its
spectral matrix obeys
\begin{equation}
  \gamma_{ab}(-\omega)=e^{-\beta\omega}\gamma_{ba}(\omega)
  \label{eq:open-system-kms-rates}
\end{equation}
for Hermitian environment operators with the Fourier convention in
\eqref{eq:open-system-bath-spectral-matrix}.  To see the sign and the order
of indices, insert a spectral resolution of \(H_E\).  If
\(H_E\ket m=E_m\ket m\), then
\[
  \gamma_{ab}(\omega)
  =
  \frac{2\pi}{Z}
  \sum_{m,n}
  e^{-\beta E_m}
  \delta\!\left(\omega-(E_n-E_m)\right)
  (E_a)_{mn}(E_b)_{nm}.
\]
The same expression for \(\gamma_{ba}(-\omega)\), with \(m\) and \(n\)
exchanged, gives
\[
  \gamma_{ba}(-\omega)
  =
  e^{-\beta\omega}\gamma_{ab}(\omega),
\]
which is equivalent to \eqref{eq:open-system-kms-rates}.  For a two-level
system with energy gap \(\Omega>0\), let
\[
  \Gamma_\downarrow=\gamma(\Omega),
  \qquad
  \Gamma_\uparrow=\gamma(-\Omega)
  =
  e^{-\beta\Omega}\Gamma_\downarrow .
\]
The population master equation is
\[
  \dot p_e=-\Gamma_\downarrow p_e+\Gamma_\uparrow p_g,
  \qquad
  p_g=1-p_e .
\]
The stationary ratio is
\[
  \frac{p_e}{p_g}
  =
  \frac{\Gamma_\uparrow}{\Gamma_\downarrow}
  =
  e^{-\beta\Omega},
\]
which is the Gibbs ratio.  This derivation is the open-system counterpart of
the detailed-balance identity in kinetic theory.

\section{Schwinger--Keldysh Influence Functional}

For a field-theoretic system variable \(\phi\) linearly coupled to an
environment operator \(\mathcal O_E\), the environment can be integrated out
on the closed time path:
\[
  e^{\ii S_{\rm IF}[\phi_+,\phi_-]}
  =
  \operatorname{Tr}_E\!\left(
  U_E[\phi_+]\rho_EU_E[\phi_-]^\dagger
  \right).
\]
Closed-time-path normalization gives
\[
  S_{\rm IF}[\phi,\phi]=0.
\]
In \(r/a\) variables,
\[
  \phi_r=\frac12(\phi_++\phi_-),
  \qquad
  \phi_a=\phi_+-\phi_-,
\]
the quadratic influence action has the form
\begin{equation}
  S_{\rm IF}^{(2)}
  =
  \int \phi_a D_R\phi_r
  +\frac{\ii}{2}\int \phi_aN\phi_a
  +O(\phi_a^3,\phi_a^2\phi_r).
  \label{eq:open-system-quadratic-influence}
\end{equation}
with
\[
  D_R(x-y)=-\ii\theta(x^0-y^0)
  \omega_E([\mathcal O_E(x),\mathcal O_E(y)]),
  \qquad
  N(x-y)=\frac12
  \omega_E(\{\mathcal O_E(x),\mathcal O_E(y)\})
\]
for a stationary environment state \(\omega_E\).  The kernel \(D_R\) is
retarded because variations of the forward and backward branches cancel for
earlier times, as in the Schwinger--Keldysh derivation of retarded response.
The kernel \(N\) is positive in the sense
\[
  \int f N f\ge0
\]
for real test functions \(f\), so that
\(\exp(\ii S_{\rm IF})\) is damped rather than amplified.  If the
environment is KMS, \(N\) and the anti-Hermitian part of \(D_R\) obey the
fluctuation--dissipation relation of
Chapter~\ref{ch:thermal-spectral-kubo-transport}.

\section{Hydrodynamic Open Systems}

Consider a single approximately conserved density \(n(t,\mathbf x)\) whose
conservation is weakly broken by a KMS environment.  The Markovian
hydrodynamic equation is
\begin{equation}
  \partial_t n+\nabla\cdot j
  =
  -\gamma(n-n_{\rm eq})+\xi .
  \label{eq:open-system-hydro-relaxation}
\end{equation}
For one homogeneous mode with susceptibility \(\chi\), the quadratic
equilibrium free energy is
\[
  F(n)=\frac{(n-n_{\rm eq})^2}{2\chi}.
\]
The Fokker--Planck equation associated with
\[
  \dot n=-\gamma(n-n_{\rm eq})+\xi,
  \qquad
  \langle\xi(t)\xi(t')\rangle=2D_n\delta(t-t')
\]
has stationary distribution proportional to \(e^{-\beta F(n)}\) precisely
when
\begin{equation}
  D_n=\gamma\chi T .
  \label{eq:open-system-einstein-noise}
\end{equation}
Indeed, the Fokker--Planck equation is
\[
  \partial_tP(n,t)
  =
  \partial_n\!\left[\gamma(n-n_{\rm eq})P(n,t)\right]
  +D_n\partial_n^2P(n,t)
  =
  -\partial_nJ_P(n,t),
\]
where
\[
  J_P(n,t)=
  -\gamma(n-n_{\rm eq})P(n,t)-D_n\partial_nP(n,t).
\]
For the equilibrium candidate
\[
  P_{\rm eq}(n)=
  Z^{-1}\exp\!\left[-\frac{(n-n_{\rm eq})^2}{2\chi T}\right],
\]
the stationary probability current vanishes for every \(n\) exactly when
\(D_n=\gamma\chi T\).
Thus
\begin{equation}
  \langle\xi(t,\mathbf x)\xi(t',\mathbf x')\rangle
  =
  2\gamma\chi T\,
  \delta(t-t')\delta^{(d)}(\mathbf x-\mathbf x')
  \label{eq:open-system-hydro-noise}
\end{equation}
in the local Markovian limit.  The relation is the Einstein relation for a
nonconserved relaxation mode; it is distinct from diffusive current noise,
where the noise enters through a divergence.

\section{Theorem Boundary}

For continuum QFT, a complete nonequilibrium construction must state the
local algebra, the reservoir states, the topology of the long-time local
limit, the domains of the current operators, and the order in which weak
coupling, thermodynamic, Markovian, hydrodynamic, and continuum limits are
taken.  For open systems, complete positivity is a statement about the
reduced dynamics of a subalgebra after tracing over an environment and taking
a specified scaling limit.  A nonlocal truncation of a closed QFT becomes an
open-system dynamics only after such an environment trace and limit have
been constructed.  For hydrodynamic open systems, the noise kernel and
relaxation matrix must be derived from the same influence functional that
produces the dissipative terms.

\begin{openproblem}[Continuum nonequilibrium states]
\label{op:continuum-nonequilibrium-steady-states}
Construct nonequilibrium steady states and open-system limits in interacting
continuum QFT with local algebras, KMS reservoirs, real-time correlators,
entropy production, hydrodynamic scaling limits, and complete-positivity
constraints derived from the microscopic dynamics.
\end{openproblem}
