\chapter{Spectral Functions, Kubo Formulae, and Transport}
\label{ch:thermal-spectral-kubo-transport}

\ConventionsRealTime

Transport coefficients are low-frequency limits of real-time response
functions.  The definitions use a thermal state, a conserved current or
stress tensor, and a source coupled to the corresponding operator.  The
derivation is linear response from the Schwinger--Keldysh generating
functional of Chapter~\ref{ch:thermal-schwinger-keldysh}.  This chapter
keeps separate three pieces that are often conflated: the commutator
spectral density, possible local contact terms in the source response, and
the order of limits needed to define a transport coefficient.

\section{Retarded Functions and Spectral Densities}

Let \(\rho_\beta\) be a KMS state at inverse temperature \(\beta\).  For
bosonic Hermitian operators \(A(t,\mathbf x)\) and \(B(t,\mathbf x)\), define
the retarded function
\begin{equation}
  G^R_{AB}(t,\mathbf x)
  =
  -\ii\,\Theta(t)\,
  \langle [A(t,\mathbf x),B(0,\mathbf 0)]\rangle_\beta
  \label{eq:kubo-retarded-definition}
\end{equation}
and the commutator spectral density
\begin{equation}
  \rho_{AB}(t,\mathbf x)
  =
  \langle [A(t,\mathbf x),B(0,\mathbf 0)]\rangle_\beta .
  \label{eq:kubo-commutator-spectral-density}
\end{equation}
Fourier transforms are defined distributionally by
\begin{equation}
  G^R_{AB}(\omega,\mathbf k)
  =
  \int dt\,d^{D-1}x\,
  e^{\ii\omega t-\ii\mathbf k\cdot\mathbf x}
  G^R_{AB}(t,\mathbf x).
  \label{eq:kubo-fourier-convention}
\end{equation}
With this convention, a retarded distribution that obeys the required
tempered-growth hypothesis is the boundary value of a function holomorphic
for \(\operatorname{Im}\omega>0\).

\paragraph{Finite-volume Lehmann representation.}
Work in a finite spatial volume.  Let \(H|n\rangle=E_n|n\rangle\), let
\(Z=\sum_n e^{-\beta E_n}<\infty\), and let \(A_{\mathbf k}\) obey
\(A_{\mathbf k}^{\dagger}=A_{-\mathbf k}\).  Then
\begin{equation}
  G^>_{AA}(\omega,\mathbf k)
  =
  \frac{2\pi}{Z}\sum_{m,n}
  e^{-\beta E_m}
  |\langle m|A_{\mathbf k}|n\rangle|^2
  \delta(\omega-E_n+E_m),
  \label{eq:kubo-greater-lehmann}
\end{equation}
and
\begin{equation}
  \rho_{AA}(\omega,\mathbf k)
  =
  \frac{2\pi}{Z}\sum_{m,n}
  e^{-\beta E_m}
  \left(1-e^{-\beta\omega}\right)
  |\langle m|A_{\mathbf k}|n\rangle|^2
  \delta(\omega-E_n+E_m).
  \label{eq:kubo-spectral-lehmann}
\end{equation}
Consequently \(\omega\,\rho_{AA}(\omega,\mathbf k)\ge0\) as a
distributional positivity statement.
Indeed, inserting a complete set of energy eigenstates into
\[
  G^>_{AA}(t,\mathbf k)
  =
  Z^{-1}\operatorname{Tr}
  \left(e^{-\beta H}A_{\mathbf k}(t)A_{-\mathbf k}(0)\right).
\]
Since
\[
  A_{\mathbf k}(t)
  =
  e^{\ii Ht}A_{\mathbf k}e^{-\ii Ht},
\]
the matrix element carries the phase
\[
  e^{\ii(E_m-E_n)t}
  \langle m|A_{\mathbf k}|n\rangle
  \langle n|A_{-\mathbf k}|m\rangle .
\]
The adjoint relation makes the product the squared modulus displayed in
\eqref{eq:kubo-greater-lehmann}, and the Fourier integral gives the delta
function \(\delta(\omega-E_n+E_m)\).

The lesser function has the same delta support with the Boltzmann weight
\(e^{-\beta E_n}\) after relabelling \(m\leftrightarrow n\).  On the support
of the delta function,
\[
  E_n-E_m=\omega,
  \qquad
  e^{-\beta E_n}=e^{-\beta\omega}e^{-\beta E_m}.
\]
Subtracting \(G^<\) from \(G^>\) gives
\eqref{eq:kubo-spectral-lehmann}.  For \(\omega>0\) every coefficient in
\eqref{eq:kubo-spectral-lehmann} is nonnegative; for \(\omega<0\) it is
nonpositive.  This is exactly \(\omega\rho_{AA}\ge0\).

\paragraph{Retarded boundary value and sign convention.}
Assume the finite-volume spectral measure above has a thermodynamic limit for
which the displayed integrals remain meaningful as tempered boundary values.
With the retarded convention \eqref{eq:kubo-retarded-definition}, the nonlocal
part of the retarded function is
\begin{equation}
  G^R_{AA}(z,\mathbf k)
  =
  \int_{\mathbb R}\frac{d\omega'}{2\pi}\,
  \frac{\rho_{AA}(\omega',\mathbf k)}{z-\omega'},
  \qquad
  \operatorname{Im}z>0,
  \label{eq:kubo-retarded-spectral-representation}
\end{equation}
up to real polynomial contact terms.  The corresponding boundary-value
convention is
\begin{equation}
  \rho_{AA}(\omega,\mathbf k)
  =
  -2\,\operatorname{Im}G^R_{AA}(\omega+\ii0,\mathbf k)
  \label{eq:kubo-rho-minus-two-im-gr}
\end{equation}
for the nonlocal part of the response.  This sign is fixed by writing the
inverse Fourier representation of \(\rho\) in
\eqref{eq:kubo-retarded-definition}: for \(\operatorname{Im}z>0\),
\[
  -\ii\int_0^\infty dt\,e^{\ii zt}e^{-\ii\omega't}
  =
  \frac{1}{z-\omega'} .
\]
This gives \eqref{eq:kubo-retarded-spectral-representation}.  Taking
\(z=\omega+\ii0\) and using
\[
  \frac{1}{\omega+\ii0-\omega'}
  =
  \operatorname{PV}\frac{1}{\omega-\omega'}
  -
  \ii\pi\delta(\omega-\omega')
\]
gives \eqref{eq:kubo-rho-minus-two-im-gr}.  Local source counterterms add
real polynomials in \(\omega\) and \(\mathbf k\) to the real-time kernel in
the absence of dissipation; they are not part of the commutator spectral
measure.

\section{KMS and Fluctuation--Dissipation}

Define the Wightman functions
\[
  G^>_{AB}(t)=\langle A(t)B(0)\rangle_\beta,\qquad
  G^<_{AB}(t)=\langle B(0)A(t)\rangle_\beta .
\]
The KMS condition gives
\[
  G^>_{AB}(t-\ii\beta)=G^<_{AB}(t)
\]
for operators in the analytic domain of the thermal state.  In frequency
space this implies detailed balance.  For \(A=B^\dagger\), the finite-volume
formula above gives
\begin{equation}
  G^<_{AA}(\omega,\mathbf k)
  =
  e^{-\beta\omega}G^>_{AA}(\omega,\mathbf k),
  \qquad
  \rho_{AA}=G^>_{AA}-G^<_{AA}.
  \label{eq:kubo-detailed-balance}
\end{equation}
Therefore the symmetrized correlator
\[
  G^{\rm sym}_{AA}(\omega,\mathbf k)
  =
  \frac12\bigl(G^>_{AA}(\omega,\mathbf k)
  +G^<_{AA}(\omega,\mathbf k)\bigr)
\]
satisfies
\begin{equation}
  G^{\rm sym}_{AA}(\omega,\mathbf k)
  =
  \frac12\coth\!\left(\frac{\beta\omega}{2}\right)
  \rho_{AA}(\omega,\mathbf k).
  \label{eq:kubo-fluctuation-dissipation}
\end{equation}
The apparent singularity at \(\omega=0\) is interpreted distributionally:
the product is meaningful when the small-\(\omega\) behavior of
\(\rho_{AA}\) is specified.  In a diffusive or viscous channel,
\(\rho_{AA}(\omega)=O(\omega)\), so the symmetrized noise has a finite
zero-frequency limit.  If a ballistic channel produces a Drude delta
function, this limit is a different observable and the dc transport
coefficient is infinite rather than a finite number.

\section{Linear Response from the Closed Time Path}

Perturb the Hamiltonian by a source \(h_B(t,\mathbf x)\) coupled to \(B\):
\begin{equation}
  H(t)=H_0-\int d^{D-1}x\,h_B(t,\mathbf x)B(t,\mathbf x).
  \label{eq:kubo-source-hamiltonian}
\end{equation}
To first order in \(h_B\), the interaction-picture evolution operator is
\[
  U_h(t,-\infty)
  =
  1+\ii\int_{-\infty}^{t}dt'\int d^{D-1}x'\,
  h_B(t',\mathbf x')B(t',\mathbf x')+O(h_B^2).
\]
Consequently
\[
  U_h(t,-\infty)^{-1}A(t,\mathbf x)U_h(t,-\infty)
  =
  A(t,\mathbf x)
  +
  \ii\int_{-\infty}^{t}dt'\,d^{D-1}x'\,
  h_B(t',\mathbf x')
  [A(t,\mathbf x),B(t',\mathbf x')]
  +O(h_B^2).
\]
Taking the thermal expectation value gives
\begin{equation}
  \delta\langle A(t,\mathbf x)\rangle
  =
  -\int dt' d^{D-1}x'\,
  G^R_{AB}(t-t',\mathbf x-\mathbf x')h_B(t',\mathbf x').
  \label{eq:kubo-linear-response}
\end{equation}
Equivalently, the source-response kernel for the convention
\(H=H_0-h_BB\) is
\[
  K^R_{AB}
  =
  \ii\Theta(t)\langle[A(t),B(0)]\rangle_\beta
  =
  -G^R_{AB}.
\]
This is the branch-difference identity obtained from the
Schwinger--Keldysh functional in
Chapter~\ref{ch:thermal-schwinger-keldysh}: differentiating with respect to
the \(a\)-type source and then setting \(J_a=0\) leaves the causal
source-response commutator.  Transport formulae below are stated in terms of
the commutator spectral density \(\rho_{AB}\), which is independent of this
source-sign convention.

\section{Contact Terms and Response Kernels}

The commutator retarded function is not always the full response kernel to a
background source.  If the operator itself depends on the source, or if the
regulated generating functional contains local source counterterms, then
\[
  \delta\langle A\rangle
  =
  (K^R_{AB}+C_{AB})\,h_B
\]
in Fourier space, where \(C_{AB}(\omega,\mathbf k)\) is local in spacetime
and therefore polynomial, or a distributional limit of polynomials, in
\(\omega,\mathbf k\).  For gauge and metric sources these terms are not
optional: they are fixed by Ward identities, equilibrium thermodynamics, and
the chosen operator scheme.

The dissipative transport coefficient is extracted from the spectral slope
of the nonlocal commutator part.  Real local contact terms can change static
susceptibilities and Euclidean zero-mode correlators, but they do not change
the positive spectral measure \(\rho_{AA}\) at nonzero frequency.  This is
why Kubo formulae below are stated using \(\rho\), with a separate warning
about Drude weights and conserved-density projections.

\section{Transport Limits and Zero-Frequency Singular Terms}

Transport is a statement about an infinite system.  In finite spatial volume
the spectrum of the Hamiltonian is discrete, so real-time correlators are
almost periodic sums of phases rather than decaying functions.  The
thermodynamic limit is therefore part of the definition of a finite
dissipative coefficient.

\begin{definition}[Transport limit datum]
\label{def:kubo-transport-limit-datum}
A transport limit datum for a pair of local operators \(A,B\) consists of
\[
  (\Sigma_L,\rho_{\beta,L},A_L,B_L,\mathcal R,\mathcal O),
\]
where \(\Sigma_L\) is a sequence of spatial regions or lattices of volume
\(V_L\), \(\rho_{\beta,L}\) is a KMS or finite Gibbs state on the regulated
system, \(A_L,B_L\) are regulated extensive or density operators with their
normalization declared, \(\mathcal R\) is the prescription for subtracting
local contact terms and conserved-density projections, and \(\mathcal O\) is
the order of limits.  For ordinary dc transport the order is
\begin{equation}
  L\to\infty,\qquad \mathbf k\to\mathbf0\ \text{in the channel specified by
  symmetry},\qquad \omega\downarrow0 .
  \label{eq:kubo-transport-order-of-limits}
\end{equation}
Changing \(\mathcal O\) can change the answer; for example static
susceptibilities and dc conductivities use different zero-frequency limits.
\end{definition}

The most elementary obstruction to a finite dc coefficient is an exactly
conserved operator that has nonzero overlap with the current.  Let \(I_i\) be
the total current in direction \(i\) in finite volume \(V\), and let
\(\bar I_i=I_i-\langle I_i\rangle_\beta\).  Suppose
\(\{Q_a\}_{a=1}^N\) are Hermitian conserved operators with zero thermal
expectation and positive covariance matrix
\begin{equation}
  C_{ab}
  =
  \frac1{2V}\langle Q_aQ_b+Q_bQ_a\rangle_\beta .
  \label{eq:kubo-conserved-covariance}
\end{equation}
Define the current--charge overlap vector
\begin{equation}
  v_{ia}
  =
  \frac1{2V}\langle \bar I_iQ_a+Q_a\bar I_i\rangle_\beta .
  \label{eq:kubo-current-charge-overlap}
\end{equation}
The projection of \(\bar I_i\) onto the span of conserved charges is
\begin{equation}
  P\bar I_i
  =
  Q_a(C^{-1})^{ab}v_{ib}.
  \label{eq:kubo-mazur-projection}
\end{equation}

\begin{lemma}[Finite-volume Mazur projection]
\label{lem:kubo-mazur-projection}
In the finite-dimensional Gibbs setting above, for every real vector \(u_i\),
the infinite-time averaged connected symmetrized current autocorrelation
satisfies
\begin{equation}
  \lim_{T\to\infty}\frac1T\int_0^T dt\,
  \frac1{2V}
  \left\langle
    \{u_i\bar I_i(t),u_j\bar I_j(0)\}
  \right\rangle_\beta
  \geq
  u_i v_{ia}(C^{-1})^{ab}v_{jb}u_j .
  \label{eq:kubo-mazur-bound}
\end{equation}
If the right hand side has a nonzero thermodynamic limit, the conductivity
has a zero-frequency singular contribution and the ordinary dc limit is not
a finite dissipative number.
\end{lemma}

\begin{proof}
Use the real Hilbert-space inner product on Hermitian operators
\[
  (X,Y)_{\beta,{\rm sym}}
  =
  \frac1{2V}\langle XY+YX\rangle_\beta .
\]
Time evolution is orthogonal for this inner product because \(\rho_\beta\)
commutes with \(H\).  For \(X=u_i\bar I_i\), let
\[
  X_\parallel
  =
  Q_a(C^{-1})^{ab}(Q_b,X)_{\beta,{\rm sym}} .
\]
Then \(X_\parallel\) is conserved and is the orthogonal projection of \(X\)
onto the conserved subspace spanned by the \(Q_a\).  The time average of the
symmetrized correlation is the squared norm of the component of \(X\) in the
zero-frequency subspace of the Liouville operator \(L=[H,\cdot]\).  This
zero-frequency subspace contains \(X_\parallel\), hence the time average is
at least
\[
  (X_\parallel,X_\parallel)_{\beta,{\rm sym}}
  =
  u_i v_{ia}(C^{-1})^{ab}v_{jb}u_j .
\]
If this quantity survives per unit volume as \(V\to\infty\), the symmetrized
current spectrum contains a term
\begin{equation}
  \frac{2\pi}{\beta}D_{ij}\delta(\omega),
  \qquad
  D_{ij}=\beta\lim_{t\to\infty}\lim_{V\to\infty}
  \frac1{2V}\langle\{\bar I_i(t),\bar I_j(0)\}\rangle_\beta ,
  \label{eq:kubo-drude-weight-definition}
\end{equation}
in the convention of \eqref{eq:kubo-fluctuation-dissipation}.  Equivalently
\(\operatorname{Re}\sigma_{ij}\) contains
\(\pi D_{ij}\delta(\omega)\).  This is a singular transport sector, not the
regular slope appearing in the finite dc Kubo formula.
\end{proof}

After the thermodynamic limit has been taken, the low-frequency real
conductivity in a fixed tensor channel must be regarded as a distribution.
The separation used below is
\begin{equation}
  \operatorname{Re}\sigma_{ij}(\omega)
  =
  \pi D_{ij}\delta(\omega)+\sigma^{\rm reg}_{ij}(\omega),
  \qquad
  \sigma^{\rm reg}_{ij}(\omega)
  =
  \frac{1}{2\omega}\rho^{\rm reg}_{J_iJ_j}(\omega,\mathbf0)
  \quad(\omega\ne0),
  \label{eq:kubo-drude-regular-decomposition}
\end{equation}
where \(D_{ij}\) is the Drude matrix produced by conserved projections and
\(\sigma^{\rm reg}_{ij}\) is locally integrable near \(\omega=0\).  The
ordinary dissipative dc coefficient is
\begin{equation}
  \sigma^{\rm dc}_{ij}
  =
  \lim_{\omega\downarrow0}\sigma^{\rm reg}_{ij}(\omega)
  =
  \lim_{\omega\downarrow0}
  \frac{1}{2\omega}\rho^{\rm reg}_{J_iJ_j}(\omega,\mathbf0),
  \label{eq:kubo-regular-dc-from-regular-spectral-slope}
\end{equation}
when this limit exists.  The singular term is a separate distributional
summand, while \(\sigma^{\rm reg}\) is the locally integrable part used in the
finite dc limit.  Figure~\ref{fig:kubo-order-of-limits} records this
order-of-limits prescription.

\begin{figure}[H]
\centering
\begin{tikzpicture}[>=Stealth,scale=1.0]
  \draw[->] (-3.3,0) -- (3.45,0) node[right] {\(\omega\)};
  \draw[->] (0,0) -- (0,2.85) node[above] {\(\operatorname{Re}\sigma(\omega)\)};
  \draw[dashed,gray] (-3.0,0.86) -- (3.0,0.86);
  \node[gray,anchor=west,fill=white,inner sep=1pt] at (2.05,0.94)
    {\(\sigma_{\rm dc}\)};
  \draw[blue!70!black,thick,smooth,domain=-3.0:3.0,samples=80]
    plot (\x,{0.28 + 1.15/(1.0+0.70*\x*\x)});
  \node[blue!70!black,anchor=west,fill=white,inner sep=1pt] at (1.05,1.62)
    {regular dissipative peak};
  \draw[orange!85!black,thick,smooth,domain=-3.0:3.0,samples=80]
    plot (\x,{0.78 + 0.055*\x*\x});
  \node[orange!85!black,anchor=west,fill=white,inner sep=1pt] at (-3.05,1.45)
    {regular channel};
  \draw[red!75!black,very thick,->] (0,0) -- (0,2.45);
  \node[red!75!black,anchor=west] at (0.18,2.35)
    {\(\pi D\,\delta(\omega)\)};
  \node[anchor=west] at (0.18,2.02)
    {singular sector};
  \draw[decorate,decoration={brace,amplitude=4pt}]
    (-0.25,0.05) -- (-0.25,0.86);
  \node[anchor=east,fill=white,inner sep=1pt] at (-0.40,0.43)
    {\(\lim_{\omega\downarrow0}\sigma^{\rm reg}\)};
  \fill[blue!70!black] (0,1.43) circle (1.7pt);
  \fill[orange!85!black] (0,0.78) circle (1.7pt);
  \node[anchor=north] at (0,-0.12) {\(0\)};
  \node[align=center,anchor=north] at (0,-0.48)
    {thermodynamic limit first; subtract \(\delta(\omega)\);
    then take the regular limit};
\end{tikzpicture}
\caption{Distributional low-frequency content of a conductivity after the
thermodynamic limit.  The finite dc coefficient is the zero-frequency limit
of the regular part \(\sigma^{\rm reg}\), equivalently the slope of
\(\rho^{\rm reg}_{JJ}(\omega,\mathbf0)\) divided by \(2\omega\).  A Drude
or superfluid weight is represented by the separate
\(\delta(\omega)\) summand.}
\label{fig:kubo-order-of-limits}
\end{figure}

\section{Projection Identities and Memory Kernels}

Kubo formulae identify transport coefficients as low-frequency response
data.  A second finite-regulator identity explains where hydrodynamic
closures enter.  Work in finite volume and let \(\mathcal A_L\) be the
finite-dimensional operator space generated by the regulated degrees of
freedom in the channel under study.  Time evolution acts on this space by
\[
  X(t)=e^{t\mathcal L}X,
  \qquad
  \mathcal L X=\ii[H_L,X],
\]
so that \(dX(t)/dt=\mathcal L X(t)\).  For linearly independent Hermitian
operators \(A_a\) with zero thermal expectation define the Kubo--Mori inner
product
\begin{equation}
  (X,Y)_\beta
  =
  \frac1\beta\int_0^\beta d\lambda\,
  \left\langle
    X^\dagger(-\ii\lambda)Y
  \right\rangle_\beta,
  \qquad
  X(-\ii\lambda)=e^{\lambda H_L}Xe^{-\lambda H_L}.
  \label{eq:kubo-mori-inner-product}
\end{equation}
Let
\[
  \chi_{ab}=(A_a,A_b)_\beta
\]
and assume that \(\chi\) is positive definite on the chosen slow subspace.
The orthogonal projection onto this subspace is
\begin{equation}
  PX=A_a(\chi^{-1})^{ab}(A_b,X)_\beta,
  \qquad
  Q=1-P .
  \label{eq:kubo-mori-projection}
\end{equation}
The projection depends on the regulated state and on the declared slow
operators.  In hydrodynamic applications \(A_a\) are Fourier modes of
conserved densities; in a transport calculation they may also include
exactly conserved charges whose overlap creates the Drude sector isolated in
Lemma~\ref{lem:kubo-mazur-projection}.

\begin{proposition}[Finite-regulator projection identity]
\label{prop:kubo-mori-zwanzig-identity}
Let \(X(t)=e^{t\mathcal L}X_0\).  On the finite operator space above,
\begin{align}
  \frac{d}{dt}PX(t)
  &=
  P\mathcal LP\,PX(t)
  +P\mathcal L e^{tQ\mathcal LQ}QX_0
  \notag\\
  &\quad
  +\int_0^t ds\,
  P\mathcal L e^{(t-s)Q\mathcal LQ}
  Q\mathcal LP\,PX(s).
  \label{eq:kubo-mori-zwanzig-projection-identity}
\end{align}
Consequently, if \(X_0\) lies in the range of \(P\), its slow coordinates
\[
  \alpha^a(t)=(\chi^{-1})^{ab}(A_b,PX(t))_\beta
\]
obey
\begin{equation}
  \dot\alpha^b(t)
  =
  \Omega_a{}^b\,\alpha^a(t)
  -\int_0^t ds\,M_a{}^b(t-s)\alpha^a(s),
  \label{eq:kubo-memory-coordinate-equation}
\end{equation}
where
\begin{align}
  \Omega_a{}^b
  &=
  (\chi^{-1})^{bc}(A_c,\mathcal L A_a)_\beta,
  \label{eq:kubo-frequency-matrix}\\
  M_a{}^b(t)
  &=
  -(\chi^{-1})^{bc}
  \left(A_c,
    \mathcal L e^{tQ\mathcal LQ}Q\mathcal L A_a
  \right)_\beta .
  \label{eq:kubo-memory-kernel-definition}
\end{align}
For a general \(X_0\), the additional term
\[
  f_X^b(t)
  =
  (\chi^{-1})^{bc}
  \left(A_c,\mathcal L e^{tQ\mathcal LQ}QX_0\right)_\beta
\]
must be added to the right hand side of
\eqref{eq:kubo-memory-coordinate-equation}.
\end{proposition}

\begin{proof}
Write \(x(t)=PX(t)\) and \(y(t)=QX(t)\).  Applying \(P\) and \(Q\) to
\(\dot X=\mathcal LX\) gives the exact block system
\[
  \dot x=P\mathcal LP\,x+P\mathcal LQ\,y,
  \qquad
  \dot y=Q\mathcal LP\,x+Q\mathcal LQ\,y .
\]
The second equation is a finite-dimensional inhomogeneous linear equation.
Duhamel's formula gives
\[
  y(t)
  =
  e^{tQ\mathcal LQ}QX_0
  +
  \int_0^t ds\,
  e^{(t-s)Q\mathcal LQ}Q\mathcal LP\,x(s).
\]
Substitution into the first block equation gives
\eqref{eq:kubo-mori-zwanzig-projection-identity}.  If
\(X_0=A_a\alpha^a(0)\), then \(QX_0=0\) and
\[
  P\mathcal LP\,PX(t)
  =
  A_b\Omega_a{}^b\alpha^a(t).
\]
The integral term is written in the same \(A_b\) basis by applying
\((\chi^{-1})^{bc}(A_c,\cdot)_\beta\), which gives
\eqref{eq:kubo-memory-coordinate-equation} with the definition
\eqref{eq:kubo-memory-kernel-definition}.  For a general initial operator the
Duhamel term involving \(QX_0\) is projected in the same way and produces
\(f_X^b(t)\).
\end{proof}

The identity is algebraic at finite regulator.  It has no decay statement in
it: the unresolved evolution \(e^{tQ\mathcal LQ}\) can be almost periodic in
finite volume, and the memory kernel can fail to have a long-time integral
before the thermodynamic limit is taken.  A hydrodynamic closure is therefore
an additional scaling statement about the kernel \(M(t)\).  In a conserved
density channel the Ward identity gives
\[
  \partial_t n_A(t,\mathbf k)=-\ii k_iJ_A^i(t,\mathbf k).
\]
If the slow variables are the densities \(n_A(\mathbf k)\), then
\(Q\mathcal L n_A(\mathbf k)\) is proportional to \(k_i\) times the part of
the current orthogonal to the density subspace.  Hence the memory term begins
at order \(k^2\):
\begin{equation}
  \partial_t n_A(t,\mathbf k)
  =
  -k^2\int_0^t ds\,
  \mathsf D_{AB}(t-s,\mathbf k)n_B(s,\mathbf k)
  +\text{reactive terms and higher-gradient corrections}.
  \label{eq:kubo-memory-diffusion-form}
\end{equation}
Replacing the kernel by a local diffusion matrix,
\[
  \int_0^t ds\,\mathsf D_{AB}(t-s,\mathbf k)n_B(s,\mathbf k)
  \leadsto
  D_{AB}n_B(t,\mathbf k),
  \qquad
  D_{AB}=\int_0^\infty dt\,\mathsf D_{AB}(t,\mathbf0),
\]
requires a thermodynamic limit, decay of the projected current kernel,
control of the \(k\to0\) limit, and prior separation of Drude sectors.  When
these hypotheses hold, the resulting matrix \(D\) is the diffusion matrix
that appears in the hydrodynamic pole
\(\omega=-\ii\lambda k^2+\cdots\) and in the Einstein relation
\(D=\Sigma\chi^{-1}\) derived in
Chapter~\ref{ch:thermal-hydrodynamics-from-ward-identities}.  If the
projected kernel has a persistent component or a nonintegrable long-time
tail, the local first-order constitutive relation is replaced by a singular
or scale-dependent low-frequency response.
Chapter~\ref{ch:thermal-hydrodynamic-fluctuations} develops the
hydrodynamic version of this infrared phenomenon.

\subsection{The Closure Datum Behind a Diffusion Pole}

The finite-regulator identity
\eqref{eq:kubo-mori-zwanzig-projection-identity} becomes a transport
statement only after a limiting datum is supplied.  Let
\(\Lambda_L\) be a sequence of spatial volumes tending to
\(\mathbb R^{d-1}\), let \(n_A^L(\mathbf k)\) be the regulated Fourier modes
of conserved densities at momenta compatible with \(\Lambda_L\), and let
\(\mathsf D^L_{AB}(t,\mathbf k)\) denote the projected kernel obtained from
\eqref{eq:kubo-memory-kernel-definition} after the Drude sector and exact
zero modes have been removed.  A local diffusive closure consists of the
following data.

\begin{hypothesis}[Projected-current hydrodynamic closure datum]
\label{hyp:projected-current-hydrodynamic-closure}
For \(\mathbf k\) in a neighborhood of \(0\) there is a matrix-valued
locally integrable function \(\mathsf D_{AB}(t,\mathbf k)\) such that:
\begin{enumerate}[label=(\roman*)]
  \item for every compact time interval \([0,T]\) and every test matrix
  \(u^{AB}(t,\mathbf k)\) smooth in \(t\) and \(\mathbf k\),
  \[
    \lim_{L\to\infty}
    \int_0^Tdt\,u^{AB}(t,\mathbf k)
    \mathsf D^L_{AB}(t,\mathbf k)
    =
    \int_0^Tdt\,u^{AB}(t,\mathbf k)
    \mathsf D_{AB}(t,\mathbf k);
  \]
  \item the Laplace transform
  \begin{equation}
    \widehat{\mathsf D}_{AB}(z,\mathbf k)
    =
    \int_0^\infty e^{-zt}\mathsf D_{AB}(t,\mathbf k)\,dt,
    \qquad \operatorname{Re}z>0,
    \label{eq:kubo-projected-memory-laplace}
  \end{equation}
  has a finite limit at \((z,\mathbf k)=(0,\mathbf0)\) after the prescribed
  subtraction of singular conserved pieces, and admits a holomorphic
  continuation to a neighborhood of this point in the \(z\)-plane with
  \[
    \widehat{\mathsf D}(z,\mathbf k)
    =
    \widehat{\mathsf D}(0,\mathbf0)+O(|z|+|\mathbf k|)
  \]
  in the finite-dimensional charge-coordinate norm;
  \item the finite-volume initial-force term
  \(P\mathcal L e^{tQ\mathcal LQ}QX_0\) has no contribution to the
  hydrodynamic pole under the same thermodynamic and small-\(\mathbf k\)
  limit, either because the initial perturbation lies in the slow subspace
  or because the projected force term decays in the limiting topology;
  \item the susceptibility matrix
  \(\chi_{AB}(\mathbf k)=
  (n_A(\mathbf k),n_B(-\mathbf k))_\beta\) has a nonsingular limit
  \(\chi_{AB}=\chi_{AB}(\mathbf0)\) on the declared hydrodynamic charge
  sector.
\end{enumerate}
\end{hypothesis}

These conditions are the theorem boundary for transport: in a concrete QFT
one must prove them, or else keep them as hypotheses.  Under
Hypothesis~\ref{hyp:projected-current-hydrodynamic-closure}, the memory
equation has a precise small-momentum meaning.  For simplicity suppress
reactive terms and write the density equation in the form
\begin{equation}
  \partial_t n_A(t,\mathbf k)
  =
  -k^2\int_0^t ds\,
  \mathsf D_{AB}(t-s,\mathbf k)n_B(s,\mathbf k)
  +O(k^3)
  \label{eq:kubo-projected-density-memory-equation}
\end{equation}
in the thermodynamic limit.

In the following displays \(A,B\) are charge-coordinate labels.  The position
of an index marks row or column placement in a matrix product and is not a
metric raising operation.  The Laplace transform of the density memory
equation is
\begin{equation}
  \left[
    z\delta_A{}^B
    +
    k^2\widehat{\mathsf D}_{A}{}^{B}(z,\mathbf k)
  \right]
  \widetilde n_B(z,\mathbf k)
  =
  n_A(0,\mathbf k)
  +O(k^3),
  \qquad \operatorname{Re}z>0 .
  \label{eq:kubo-memory-laplace-density-equation}
\end{equation}
If \(D_A{}^B=\widehat{\mathsf D}_{A}{}^{B}(0,\mathbf0)\) is diagonalizable
on the charge sector and \(\lambda_r\) is one of its eigenvalues, analytic
perturbation of the finite matrix
\[
  z\mathbf 1+k^2\widehat{\mathsf D}(z,\mathbf k)
\]
gives a pole of the density response at
\begin{equation}
  z_r(\mathbf k)
  =
  -k^2\lambda_r+o(k^2).
  \label{eq:kubo-memory-diffusive-laplace-pole}
\end{equation}
With the real-time Fourier convention of
\eqref{eq:kubo-fourier-convention}, \(z=-\ii\omega\), this is the
hydrodynamic pole
\begin{equation}
  \omega_r(\mathbf k)
  =
  -\ii\lambda_r k^2+o(k^2).
  \label{eq:kubo-memory-diffusive-frequency-pole}
\end{equation}
The construction of the diffusion pole therefore has three distinct steps:
construct the thermodynamic projected-current kernel, prove the
zero-frequency regularity of its Laplace transform, and then expand the
resolvent of the closed slow-coordinate equation.

The conductivity matrix entering the Einstein relation is the same datum
expressed in current variables.  Let \(\Sigma_{AB}\) denote the regular
zero-frequency limit of the current response after subtracting the Drude
sector and contact terms.  In a purely diffusive charge channel,
\begin{equation}
  \Sigma_{AB}
  =
  D_A{}^C\chi_{CB}.
  \label{eq:kubo-memory-einstein-relation}
\end{equation}
This relation follows from the continuity equation: an external static
source changes the equilibrium charge density by
\(\delta n_A=\chi_{AB}\delta\mu^B\), while the hydrodynamic constitutive
response to the slowly varying source is
\[
  J_A^i=-D_A{}^B\partial_i\delta n_B
  =-\Sigma_{AB}\partial_i\delta\mu^B .
\]
Equation~\eqref{eq:kubo-memory-einstein-relation} is therefore a statement
about the same limiting kernel seen in two coordinate systems, not an
independent postulate.

\section{Conductivity}

For a conserved current \(J^\mu\) coupled to a background gauge field
\(A_\mu^{\rm ext}\), the electric field in the Fourier convention
\eqref{eq:kubo-fourier-convention} is
\[
  E_i(\omega,\mathbf k)=\ii\omega A_i^{\rm ext}(\omega,\mathbf k)
  -\ii k_i A_0^{\rm ext}(\omega,\mathbf k).
\]
At zero spatial momentum and in a gauge with \(A_0^{\rm ext}=0\), define the
full transverse response kernel \(K_{ij}\) by the conventional
background-gauge-field sign
\[
  \langle J_i(\omega)\rangle
  =
  -K_{ij}(\omega,\mathbf0)A_j^{\rm ext}(\omega)
  =
  \sigma_{ij}(\omega)E_j(\omega).
\]
Thus
\[
  \sigma_{ij}(\omega)
  =
  -\frac{1}{\ii\omega}K_{ij}(\omega,\mathbf0),
  \qquad
  K_{ij}=G^R_{J_iJ_j}+C_{ij}.
\]
For \(\omega\ne0\), the dissipative real part is determined by the spectral
density:
\begin{equation}
  \operatorname{Re}\sigma_{ij}(\omega)
  =
  \frac{1}{2\omega}\,
  \rho_{J_iJ_j}(\omega,\mathbf0)
  \quad
  \text{in a transverse channel after contact terms are separated.}
  \label{eq:kubo-conductivity-spectral-slope}
\end{equation}
The dc conductivity, when finite, is therefore
\[
  \sigma_{ij}^{\rm dc}
  =
  \lim_{\omega\downarrow0}
  \frac{1}{2\omega}\rho_{J_iJ_j}(\omega,\mathbf0),
\]
with the thermodynamic limit taken before the \(\omega\to0\) limit.  If
momentum conservation or another exactly conserved quantity overlaps with
the current, a \(\delta(\omega)\) Drude contribution is present and the dc
conductivity is infinite unless the conserved contribution is relaxed or
projected out.

\section{Shear and Bulk Viscosity}

Couple the theory to a spatial metric perturbation
\[
  g_{ij}(t,\mathbf x)=\delta_{ij}+h_{ij}(t,\mathbf x).
\]
The stress tensor is defined by the first variation of the regulated
generating functional with respect to the metric source.  In an isotropic
thermal state, take a shear perturbation \(h_{xy}(t)\) at zero spatial
momentum.  The dissipative part of the stress response is characterized by
the shear viscosity \(\eta\).  With the retarded sign convention
\eqref{eq:kubo-retarded-definition} and the spectral convention
\eqref{eq:kubo-rho-minus-two-im-gr},
\begin{equation}
  \eta
  =
  \lim_{\omega\downarrow0}
  \frac{1}{2\omega}\,
  \rho_{T_{xy}T_{xy}}(\omega,\mathbf0)
  =
  -\lim_{\omega\downarrow0}
  \frac{1}{\omega}\,
  \operatorname{Im}G^R_{T_{xy}T_{xy}}(\omega,\mathbf0).
  \label{eq:kubo-shear-viscosity-sign-correct}
\end{equation}
The equality between the two expressions is exactly
\eqref{eq:kubo-rho-minus-two-im-gr}.  Positivity of
\(\omega\rho_{T_{xy}T_{xy}}\) gives \(\eta\ge0\) when the limit exists and
no singular zero-frequency contribution is present.

The bulk channel requires a projection.  For a neutral relativistic fluid in
\(d=D-1\) spatial dimensions, define
\[
  \mathcal P
  =
  \frac{1}{d}T^i_{\ i}-c_s^2 T^{00},
  \qquad
  c_s^2=\left(\frac{\partial p}{\partial\varepsilon}\right)_{\rm eq}.
\]
The subtraction removes the overlap of the pressure fluctuation with the
conserved energy density.  In charged fluids one must project out all
conserved densities using the susceptibility matrix.  After this projection,
the bulk viscosity is
\[
  \zeta
  =
  \lim_{\omega\downarrow0}
  \frac{1}{2\omega}\rho_{\mathcal P\mathcal P}(\omega,\mathbf0),
\]
again with the thermodynamic and zero-momentum limits taken before the
zero-frequency limit.

\section{Dispersion Relations and Sum-Rule Boundaries}

Retardedness gives analyticity in the upper half \(\omega\)-plane.  If the
response kernel is sufficiently bounded after subtracting its local
polynomial terms, then Cauchy's theorem gives a subtracted dispersion
relation.  A typical once-subtracted form is
\begin{equation}
  G^R(\omega)-G^R(0)
  =
  \omega
  \int_{\mathbb R}\frac{d\omega'}{2\pi}\,
  \frac{\rho(\omega')}
       {\omega'(\omega-\omega'+\ii0)} .
  \label{eq:kubo-once-subtracted-dispersion}
\end{equation}
This formula is not a definition.  It follows only after the large semicircle
is controlled and after all contact terms have been declared.  Sum rules for
conductivity or viscosity compare the small-\(\omega\) transport slope with
Euclidean or equal-time data only under these additional analytic and
ultraviolet assumptions.

\section{Hydrodynamic Pole Boundary}

Kubo formulae define transport coefficients when the indicated limits exist
and commute with the thermodynamic and zero-momentum limits in the stated
order.  Hydrodynamics begins after these coefficients are inserted into the
long-wavelength Ward identities for conserved densities.  The microscopic
Kubo formula by itself does not prove local equilibration, the derivative
expansion, or the dominance of hydrodynamic poles; those are additional
dynamical statements developed in the following chapters.

\begin{openproblem}[Rigorous Kubo-to-hydrodynamics bridge]
\label{op:kubo-hydrodynamics-bridge}
Give sufficient QFT hypotheses under which retarded stress-tensor and current
correlators have the analyticity, spectral positivity, clustering, and
low-frequency limits required by the Kubo formulae, and prove that the same
coefficients control the long-time hydrodynamic effective theory.
\end{openproblem}
