\chapter{Kinetic Theory As A Controlled Limit}
\label{ch:thermal-kinetic-theory-controlled-limit}

\ConventionsMixed

Kinetic theory is the long-wavelength, weakly off-equilibrium description in
which selected real-time QFT correlators are represented by on-shell
phase-space densities.  The construction has three inputs: a quasiparticle
spectral regime, an on-shell collision kernel extracted from microscopic
scattering or finite-temperature amplitudes, and a hierarchy separating the
microscopic correlation scale from the spacetime scale of variation.  This
chapter develops the relativistic Boltzmann framework with the assumptions
kept explicit, derives detailed balance and the entropy inequality, and
records the precise point at which kinetic theory becomes hydrodynamics.

Let spacetime dimension be \(D=d+1\), with spatial momentum
\(\mathbf p\in\mathbb R^d\).  For species \(a\),
\[
  p^\mu=(E_a(\mathbf p),\mathbf p),
  \qquad
  E_a(\mathbf p)=\sqrt{\mathbf p^2+m_a^2}.
\]
The Lorentz-invariant on-shell measure used below is
\begin{equation}
  d\Pi_a(p)
  =
  \frac{d^d\mathbf p}{(2\pi)^d\,2E_a(\mathbf p)}.
  \label{eq:kinetic-onshell-measure}
\end{equation}
The factor \(2p^\mu d\Pi_a\) has the standard equal-time interpretation:
\[
  2E_a\,d\Pi_a=\frac{d^d\mathbf p}{(2\pi)^d}.
\]
Internal degeneracies may be included by replacing the species label \(a\)
with a refined label, or by inserting an explicit degeneracy \(g_a\).

\section{Quasiparticle Distribution From Spectral Data}

Let \(G^<_a(x,y)\) be a Wightman two-point function for a microscopic field
or for a gauge-invariant quasiparticle interpolating operator in a regulated
theory.  Introduce center and relative coordinates
\[
  X=\frac{x+y}{2},\qquad s=x-y,
\]
and define the Wigner transform
\[
  G^<_a(X,p)=\int d^Ds\,e^{\ii p\cdot s}G^<_a(X+s/2,X-s/2).
\]
The quasiparticle distribution \(f_a(X,\mathbf p)\) is the coefficient of
the positive-energy pole contribution when the spectral density has the form
\begin{equation}
  \rho_a(X,p)
  =
  2\pi Z_a(X,\mathbf p)\operatorname{sgn}(p^0)
  \delta\!\left((p^0)^2-E_a(\mathbf p)^2\right)
  +\rho^{\rm off}_a(X,p),
  \label{eq:kinetic-narrow-spectral-form}
\end{equation}
with the off-shell remainder small in the topology used for the gradient
expansion.  The invariant delta function satisfies
\[
  \delta\!\left((p^0)^2-E_a(\mathbf p)^2\right)
  =
  \frac{1}{2E_a(\mathbf p)}
  \left[\delta(p^0-E_a(\mathbf p))
  +\delta(p^0+E_a(\mathbf p))\right],
\]
so, near \(p^0=E_a(\mathbf p)\),
\[
  G^<_a(X,p)
  =
  \frac{2\pi Z_a(X,\mathbf p)}{2E_a(\mathbf p)}
  f_a(X,\mathbf p)
  \delta(p^0-E_a(\mathbf p))
  +\text{controlled off-shell remainder}.
\]
The factor \(1/(2E_a)\) is the mass-shell Jacobian.  Dropping it is a
change of normalization from the invariant spectral residue to an
equal-time density, and it must then be accompanied by compensating changes
in the pole residue and in the collision projector.
The pole approximation requires the spectral width to be small compared with
the quasiparticle energy,
\begin{equation}
  \varepsilon_{\rm width}(\mathbf p)
  =
  \frac{\Gamma_a(\mathbf p)}{E_a(\mathbf p)}
  \ll1.
  \label{eq:kinetic-width-small-parameter}
\end{equation}
It does not require one universal inequality involving
\(\Gamma_a\ell_{\rm variation}\).  Let
\[
  \tau_{\rm qp}\sim E_a^{-1},\qquad
  \tau_{\rm coll}\sim\Gamma_a^{-1},\qquad
  \tau_{\rm mem},\qquad
  \tau_X
\]
denote, respectively, the microscopic oscillation time, the damping or
mean-free time, the memory time of the connected self-energy kernel, and the
macroscopic time scale on which \(f_a(X,\mathbf p)\) varies.  The Markov
gradient expansion requires
\begin{equation}
  \varepsilon_{\rm grad}
  =
  \frac{\tau_{\rm mem}}{\tau_X}\ll1,
  \label{eq:kinetic-memory-gradient-small-parameter}
\end{equation}
with the analogous spatial statement using the memory length and the
macroscopic gradient length.  The relation between \(\tau_X\) and
\(\tau_{\rm coll}\) distinguishes regimes:
\[
\begin{array}{lll}
\text{collisionless kinetic window} &
  \tau_{\rm mem}\ll \tau_X\ll\tau_{\rm coll},
  & \Gamma_a\tau_X\ll1,\\[2pt]
\text{kinetic relaxation window} &
  \tau_{\rm mem}\ll \tau_X\sim\tau_{\rm coll},
  & \Gamma_a\tau_X=O(1),\\[2pt]
\text{hydrodynamic window} &
  \tau_{\rm mem}\ll\tau_{\rm coll}\ll\tau_X,
  & \Gamma_a\tau_X\gg1 .
\end{array}
\]
Thus a Boltzmann equation can describe relaxation on the collision time, while
hydrodynamics is the still longer scale on which nonconserved kinetic modes
have already decayed.  These inequalities are hypotheses on the state,
coupling regime, and initial correlations; they are not consequences of
conservation laws alone.

\section{Quasiparticle Projection of the Two-Point Equation}
\label{sec:kinetic-quasiparticle-projection-two-point-equation}

The Boltzmann drift term is the on-shell projection of the leading
Schwinger--Keldysh two-point equation.  We record the projection explicitly
because it is the point where an operator statement about correlators becomes
a phase-space equation.

\begin{controlledapproximation}[Force-free quasiparticle projection]
\label{ca:kinetic-force-free-quasiparticle-projection}
Consider one scalar species in flat spacetime, with constant mass and a
single narrow positive-energy peak.  Let
\begin{equation}
  F(p)=(p^0)^2-E(\mathbf p)^2,\qquad
  E(\mathbf p)=\sqrt{\mathbf p^2+m^2},
  \label{eq:kinetic-free-mass-shell-function}
\end{equation}
and assume that the leading-gradient Kadanoff--Baym equation has the form
\begin{equation}
  \frac12\{F,G^<\}_{\rm kin}=I^<[G],
  \label{eq:kinetic-leading-kb-force-free}
\end{equation}
where \(I^<\) is the collision functional and
\begin{equation}
  \{A,B\}_{\rm kin}
  =
  \partial_{p^0}A\,\partial_tB
  -\partial_tA\,\partial_{p^0}B
  -\partial_{\mathbf p}A\cdot\nabla_{\mathbf x}B
  +\nabla_{\mathbf x}A\cdot\nabla_{\mathbf p}B .
  \label{eq:kinetic-poisson-bracket-convention}
\end{equation}
Use the positive-energy ansatz
\begin{equation}
  G^<(X,p)=
  \frac{2\pi Z(\mathbf p)}{2E(\mathbf p)}
  f(X,\mathbf p)\delta(p^0-E(\mathbf p))
  \label{eq:kinetic-positive-energy-lesser-ansatz}
\end{equation}
with \(Z\) independent of \(X\) at this order.  Terms from mean-field
forces, Berry curvature, spatially varying masses, and off-shell memory
kernels are omitted in this controlled approximation; each of them modifies
the left-hand side by additional explicitly computable phase-space terms.
\end{controlledapproximation}

\begin{remark}[Projection of the drift term]
\label{rem:kinetic-quasiparticle-drift-projection}
Under
Controlled Approximation~\ref{ca:kinetic-force-free-quasiparticle-projection},
the positive-energy shell projection of
\eqref{eq:kinetic-leading-kb-force-free} is
\begin{equation}
  p^\mu\partial_\mu f(X,\mathbf p)=\mathcal C[f](X,\mathbf p),
  \qquad
  p^\mu=(E(\mathbf p),\mathbf p),
  \label{eq:kinetic-projected-boltzmann-equation}
\end{equation}
with
\begin{equation}
  \mathcal C[f](X,\mathbf p)
  =
  \frac{2E(\mathbf p)}{Z(\mathbf p)}
  \int_0^\infty\frac{dp^0}{2\pi}\,I^<[G](X,p).
  \label{eq:kinetic-projected-collision-functional}
\end{equation}

Since \(F\) is independent of \(X\),
\begin{equation}
  \{F,G^<\}_{\rm kin}
  =
  2p^0\partial_tG^<
  +2\mathbf p\cdot\nabla_{\mathbf x}G^< .
  \label{eq:kinetic-free-bracket}
\end{equation}
Using \eqref{eq:kinetic-positive-energy-lesser-ansatz} and the
\(X\)-independence of \(Z\) and \(E\),
\begin{equation}
  \frac12\{F,G^<\}_{\rm kin}
  =
  \frac{2\pi Z(\mathbf p)}{2E(\mathbf p)}
  \bigl(p^0\partial_t+\mathbf p\cdot\nabla_{\mathbf x}\bigr)
  f(X,\mathbf p)\,\delta(p^0-E(\mathbf p)).
  \label{eq:kinetic-projected-bracket-before-integral}
\end{equation}
Integrating over \(p^0>0\) gives
\[
  \frac{Z(\mathbf p)}{2E(\mathbf p)}
  \left(E(\mathbf p)\partial_tf+\mathbf p\cdot\nabla_{\mathbf x}f\right)
  =
  \int_0^\infty\frac{dp^0}{2\pi}I^<[G](X,p).
\]
Multiplying by \(2E(\mathbf p)/Z(\mathbf p)\) gives
\begin{equation}
  E(\mathbf p)\partial_tf+\mathbf p\cdot\nabla_{\mathbf x}f
  =
  \frac{2E(\mathbf p)}{Z(\mathbf p)}
  \int_0^\infty\frac{dp^0}{2\pi}I^<[G](X,p),
  \label{eq:kinetic-projected-bracket-after-integral}
\end{equation}
which is \eqref{eq:kinetic-projected-boltzmann-equation}.  Dividing by
\(E\) gives the coordinate-time form
\((\partial_t+\mathbf v\cdot\nabla_{\mathbf x})f=C^{(t)}[f]\) with
\(\mathbf v=\mathbf p/E\).
\end{remark}

\section{Microscopic Scalar Derivation of a Collision Kernel}
\label{sec:kinetic-scalar-sk-collision-kernel}

We now derive one collision kernel rather than only declaring it.  The example
is deliberately modest: a weakly coupled real scalar with interaction
\(-\lambda\phi^4/4!\), no gauge fields, no collinear singularities, and a
single narrow quasiparticle branch.  Gauge-theory soft exchange, screening,
collinear splitting, and LPM resummation are separate layers, not consequences
of this scalar calculation.

\begin{controlledapproximation}[Weak scalar 2PI/Kadanoff--Baym regime]
\label{ca:kinetic-scalar-2pi-kb-regime}
Fix a real scalar field in a regulated Schwinger--Keldysh theory with
renormalized local interaction
\[
  \mathcal L_{\rm int}=-\frac{\lambda_R}{4!}\phi^4 .
\]
The local counterterm datum at this order consists of
\[
  \frac{\delta Z}{2}\partial_\mu\phi\,\partial^\mu\phi
  -\frac{\delta m^2}{2}\phi^2
  -\frac{\delta\lambda}{4!}\phi^4 ,
\]
with renormalization conditions fixing the quasiparticle pole mass, unit
residue at the chosen reference point, and the four-point kernel defining
\(\lambda_R\).  In a truncated 2PI equation the local counterterms used in the
one-point/two-point equation and in the four-point Bethe--Salpeter kernel are
part of the same finite chart.  We assume such a chart has been fixed at the
regulator scale before taking the kinetic limit.
Use the two-particle-irreducible truncation that retains the sunset functional
\begin{equation}
  \Phi_2[G]
  =
  -\frac{\lambda_R^2}{48}
  \int_{\mathcal C}\dd^Dx\,\dd^Dy\,
  G(x,y)^4,
  \label{eq:kinetic-scalar-2pi-sunset-functional}
\end{equation}
besides the local tadpole and counterterm pieces.  The local tadpole shifts
the quasiparticle mass \(m_R\), while \(\delta Z\) and the local 2PI
counterterms fix the dispersion and four-point renormalization conditions;
none of these local terms produces a \(2\to2\) collision kernel.
The nonlocal self-energy obtained by differentiating
\(\Phi_2\) has the sunset form
\begin{equation}
  \Sigma_{\mathcal C}(x,y)
  =
  -\frac{\ii\lambda_R^2}{6}\,G_{\mathcal C}(x,y)^3
  +\Sigma_{\rm ct}(x,y),
  \label{eq:kinetic-scalar-sunset-self-energy}
\end{equation}
where \(\Sigma_{\rm ct}\) is local in the renormalized mass, field-strength,
coupling, and 2PI-truncation chart.
Assume:
\begin{enumerate}
\item the connected four-point initial correlations either vanish in the
      prepared state or contribute a residual
      \(\varepsilon_{\rm corr}\) after the microscopic transient;
\item the memory kernel built from the connected self-energy decays in
      relative time on a scale \(\tau_{\rm mem}\);
\item the spectral function has the narrow positive-energy form
      \eqref{eq:kinetic-narrow-spectral-form} with
      \(\Gamma/E\ll1\);
\item the Wigner transform varies on a scale \(\tau_X\) with
      \(\tau_{\rm mem}/\tau_X\ll1\).
\end{enumerate}
\end{controlledapproximation}

Under Controlled
Approximation~\ref{ca:kinetic-scalar-2pi-kb-regime}, the lesser/greater
Kadanoff--Baym equation has, to first Wigner-gradient order,
\begin{equation}
  \frac12
  \{p^2-m_R^2-\operatorname{Re}\Sigma_R,\,
    G^<\}_{\rm kin}
  =
  \frac12\left(\Sigma^<G^>-\Sigma^>G^<\right)
  +R_{\rm grad}+R_{\rm mem}.
  \label{eq:kinetic-scalar-kb-with-collision}
\end{equation}
The exact closed-time-path equation before these approximations is still a
reversible correlator equation.  Irreversible kinetics enters only after the
quasiparticle projection, first-gradient/Markov truncation, closure of higher
correlators, decay of initial correlations, and kinetic scaling limit have
been chosen.  In equilibrium the KMS relation makes the gain/loss difference
vanish.  Away from equilibrium the right side is the projected difference
between the two cut orientations of the sunset graph.

\paragraph{Cut sunset and on-shell rates.}
For the cut lines it is safer to display the conventional factors of \(\ii\).
Write
\[
  \mathcal G^{<,>}:=\ii G^{<,>},\qquad
  \mathcal \Sigma^{<,>}:=\ii\Sigma^{<,>};
\]
these are the real Wigner-space cut functions.  The real-scalar on-shell
ansatz contains both energy signs:
\begin{align}
  \mathcal G^<(X,p)
  &=
  2\pi\delta(p^2-m_R^2)
  \Big[
    \theta(p^0)f(X,\mathbf p)
    +
    \theta(-p^0)\bigl(1+f(X,-\mathbf p)\bigr)
  \Big]
  +\mathcal G^<_{\rm off},
  \nonumber\\
  \mathcal G^>(X,p)
  &=
  2\pi\delta(p^2-m_R^2)
  \Big[
    \theta(p^0)\bigl(1+f(X,\mathbf p)\bigr)
    +
    \theta(-p^0)f(X,-\mathbf p)
  \Big]
  +\mathcal G^>_{\rm off}.
  \label{eq:kinetic-scalar-lesser-greater-on-shell}
\end{align}
The positive-energy branch of \(2\pi\delta(p^2-m_R^2)\) integrates to
\(1/(2E)\), so the displayed scalar ansatz is the \(Z=1\) instance of the
invariant pole convention above, not an equal-time convention with the
Jacobian absorbed into \(f\).
Thus, for \(p_i=(E_i,\mathbf p_i)\),
\[
  \mathcal G^<(p_i)\leadsto f_i,\qquad
  \mathcal G^>(p_i)\leadsto 1+f_i,\qquad
  \mathcal G^<(-p_i)\leadsto 1+f_i,\qquad
  \mathcal G^>(-p_i)\leadsto f_i ,
\]
with the common on-shell measure \(d\Pi_i\) suppressed in this mnemonic.
The real-field relation is
\begin{equation}
  \mathcal G^<(-p)=\mathcal G^>(p),
  \qquad
  \mathcal G^>(-p)=\mathcal G^<(p),
  \label{eq:kinetic-real-scalar-wightman-reversal}
\end{equation}
which is why the same distribution \(f\) appears on both energy branches.

The nonlocal sunset prefactor in
\eqref{eq:kinetic-scalar-sunset-self-energy} is \(1/3!\).  For the crossed
\(2\to2\) cut at external \(p_1^0=E_1>0\), one of the three internal cut lines
is put on the negative-energy branch and then reversed to become the incoming
partner \(p_2\), while the other two are the outgoing momenta \(p_3,p_4\).
For example,
\[
  (k_1,k_2,k_3)=(-p_2,p_3,p_4)
  \quad\Longrightarrow\quad
  p_1-k_1-k_2-k_3=0
  \ \Longleftrightarrow\
  p_1+p_2=p_3+p_4 .
\]
The two other choices of the negative-energy line turn the \(1/3!\) sunset
factor into \(3/3!=1/2\).  The exchange \(p_3\leftrightarrow p_4\) is not a
fourth Wick-contraction factor; it is the same final state represented twice
by the ordered full-product measure \(d\Pi_3d\Pi_4\).  With the tree amplitude
convention
\(|\mathcal M_{\phi\phi\to\phi\phi}|^2=\lambda_R^2\), the physical-channel
cut self-energies in this full-product convention are therefore
\begin{align}
  \mathcal\Sigma^<(X,p_1)
  &=
  \frac{\lambda_R^2}{2}
  \int d\Pi_2d\Pi_3d\Pi_4\,
  (2\pi)^D\delta^{(D)}(p_1+p_2-p_3-p_4)
  \nonumber\\
  &\qquad\qquad\times
  f_3f_4(1+f_2)+R_{\Sigma<},
  \label{eq:kinetic-scalar-sigma-lesser-cut}\\
  \mathcal\Sigma^>(X,p_1)
  &=
  \frac{\lambda_R^2}{2}
  \int d\Pi_2d\Pi_3d\Pi_4\,
  (2\pi)^D\delta^{(D)}(p_1+p_2-p_3-p_4)
  \nonumber\\
  &\qquad\qquad\times
  f_2(1+f_3)(1+f_4)+R_{\Sigma>}.
  \label{eq:kinetic-scalar-sigma-greater-cut}
\end{align}
The lesser cut is the gain self-energy for an external particle \(p_1\):
the negative-energy branch supplies the Bose factor \(1+f_2\) for the crossed
incoming partner, and the two positive-energy branches supply \(f_3f_4\).
The greater cut is the loss self-energy: the crossed partner gives \(f_2\),
while the outgoing branches give \((1+f_3)(1+f_4)\).  A positive-frequency-only
ansatz would only produce all-positive cut momenta; it would not produce the
crossed conservation law \(p_1+p_2=p_3+p_4\) or the two different gain/loss
assignments.

The displayed self-energy coefficient is the physical-channel tree weight
with the identical-final-state divisor already included.  If the same phase
space is instead implemented as a quotient by \(p_3\leftrightarrow p_4\), the
finite divisor moves from the coefficient into the measure.  The product
measure, the channel coefficient, and the later entropy symmetrization must
be chosen as a single convention; the Bose factors do not repair a mismatch.
The residuals \(R_{\Sigma<,>}\) collect off-shell spectral tails, higher
gradients, memory corrections, and connected correlation corrections not
represented by the product of two-point functions.

Using
\[
  \mathcal G^<(p_1)=2\pi\delta_+(p_1^2-m_R^2)f_1,\qquad
  \mathcal G^>(p_1)=2\pi\delta_+(p_1^2-m_R^2)(1+f_1),
\]
and the cut-normalized convention
\(\mathcal I^<=(\mathcal\Sigma^<\mathcal G^>
-\mathcal\Sigma^>\mathcal G^<)/2\), projection of
\eqref{eq:kinetic-scalar-kb-with-collision} on the positive mass shell gives
\begin{align}
  p_1^\mu\partial_\mu f_1
  &=
  \frac{\lambda_R^2}{4}
  \int d\Pi_2d\Pi_3d\Pi_4\,
  (2\pi)^D\delta^{(D)}(p_1+p_2-p_3-p_4)
  \nonumber\\
  &\quad\times
  \Big[
  f_3f_4(1+f_1)(1+f_2)
  -
  f_1f_2(1+f_3)(1+f_4)
  \Big]
  +R_{\rm kin}.
  \label{eq:kinetic-scalar-derived-boltzmann-kernel}
\end{align}
The displayed \(\lambda_R^2/4\) is the product of the sunset
self-energy weight \(\lambda_R^2/2\) and the Kadanoff--Baym collision
projection factor \(1/2\).  After division by \(E_1\), the coordinate-time
prefactor is \(\lambda_R^2/(4E_1)\) in the same full-product convention.
Changing to a quotient convention for identical final states moves the same
finite factor into the transition measure; it must not be reintroduced
through the sign assignment or through an additional ``outgoing
permutation'' multiplier.

\paragraph{Memory, closure, and secular terms.}
The Markov step is a Taylor expansion of a memory integral, not the
replacement of \(\tau_{\rm mem}\) by the damping time.  If
\[
  \int_0^\infty\dd s\, |K(s)| <\infty,\qquad
  \tau_{\rm mem}
  =
  \frac{\int_0^\infty\dd s\,s|K(s)|}{\int_0^\infty\dd s\,|K(s)|},
\]
then for a component of \(f\) varying on \(\tau_X\),
\begin{equation}
  \Delta_K[f]
  :=
  \int_0^\infty\dd s\,K(s)f(X^0-s,\mathbf X,\mathbf p)
  -
  \left(\int_0^\infty\dd s\,K(s)\right)f(X,\mathbf p)
  \label{eq:kinetic-markov-memory-difference}
\end{equation}
satisfies the direct estimate
\begin{equation}
  |\Delta_K[f]|
  \le
  \sup_{r\ge0}|\partial_{X^0}f(X^0-r,\mathbf X,\mathbf p)|
  \int_0^\infty\dd s\,s|K(s)|.
  \label{eq:kinetic-markov-memory-expansion}
\end{equation}
Only after declaring a size \(F_X\) for \(f\), a derivative bound
\(|\partial_{X^0}f|\le F_X/\tau_X\), and a noncancellation condition
\[
  \left|\int_0^\infty K(s)\,\dd s\right|
  \ge
  c_K\int_0^\infty |K(s)|\,\dd s
  \qquad(c_K>0)
\]
does this become the relative statement
\[
  \frac{|\Delta_K[f]|}
       {F_X\left|\int_0^\infty K(s)\,\dd s\right|}
  \le
  c_K^{-1}\frac{\tau_{\rm mem}}{\tau_X}.
\]
Without the normalization and noncancellation hypotheses,
\(O(\tau_{\rm mem}/\tau_X)\) is not a dimensionally meaningful replacement
for the memory integral.
The closure assumption is similarly explicit: connected higher correlators and
initial correlations are either included in \(R_{\rm kin}\) or shown to decay
after the microscopic transient.  This is the QFT version of molecular chaos.

The same calculation explains why the kinetic time is nontrivial.  A naive
fixed-order perturbation theory contains retarded-advanced products of the
form
\begin{equation}
  G_R(p)G_A(p)
  \simeq
  \frac{1}{(p^2-m_R^2)^2+(p^0\Gamma)^2}
  \longrightarrow
  \frac{\pi}{p^0\Gamma}\delta(p^2-m_R^2)
  \label{eq:kinetic-pinch-enhancement}
\end{equation}
in the narrow-width limit.  Since \(\Gamma=O(\lambda_R^2)\) in the scalar
example, the apparent higher-order sunset correction is enhanced by
\(1/\Gamma\) on times \(t\sim\Gamma^{-1}\).  The Boltzmann equation resums
these secularly enhanced insertions into time evolution of \(f\).

\paragraph{Remainder topology.}
A concrete kinetic claim must say where the residual is small.  One convenient
norm for scalar distributions is the weighted phase-space \(L^1\) norm
\[
  \|h\|_{w}
  =
  \int 2d\Pi(p)\,w(\mathbf p)|h(\mathbf p)|,
  \qquad
  w(\mathbf p)=1+E(\mathbf p)^2 .
\]
The scalar derivation records the bound below only as a
controlled-approximation target.  To make it a theorem
one must prove the collision map is locally Lipschitz from an \(L^1_{w'}\)
moment ball to \(L^1_w\), the memory and off-shell kernels obey the weighted
moment estimates uniformly for \(0\le X^0\le T_{\rm kin}\), and the connected
correlation residual is bounded in the same norm.  With those hypotheses
declared, the desired residual statement is
\begin{equation}
  \|R_{\rm kin}\|_{w}
  \le
  C_w\Gamma
  \left(
    \varepsilon_{\rm grad}
    +
    \varepsilon_{\rm width}
    +
    \varepsilon_{\rm off}
    +
    \varepsilon_{\rm corr}
  \right)
  (1+\|f\|_{w'}),
  \label{eq:kinetic-scalar-remainder-ledger}
\end{equation}
where \(\varepsilon_{\rm off}\) is the weighted off-shell spectral tail,
\(\varepsilon_{\rm grad}\) includes the normalized memory estimate above,
\(\varepsilon_{\rm corr}\) bounds the connected initial-correlation and
closure remainders after the microscopic transient, and \(w'\) is a weight
large enough to control the moments in the collision integral.  The formula is
a ledger for the controlled approximation, not an established bound in this
chapter: setting one of these terms to zero is an additional theorem or
assumption, not a definition of kinetic theory.

\section{Boltzmann Equation and Collision Kernel}

The covariant Boltzmann equation is
\begin{equation}
  p^\mu\partial_\mu f_a(X,\mathbf p)=\mathcal C_a[f](X,\mathbf p).
  \label{eq:kinetic-covariant-boltzmann}
\end{equation}
In coordinate time this is
\[
  \left(\partial_t+\mathbf v_a\cdot\nabla_{\mathbf x}\right)f_a
  =
  C^{(t)}_a[f],
  \qquad
  \mathbf v_a=\nabla_{\mathbf p}E_a,
  \qquad
  C^{(t)}_a=\frac{\mathcal C_a}{E_a}.
\]

For \(a(p_1)b(p_2)\to c(p_3)d(p_4)\), define
\[
  \eta_a=
  \begin{cases}
    +1,& a\ \text{bosonic},\\
    -1,& a\ \text{fermionic}.
  \end{cases}
\]
For each physical channel choose a nonnegative transition weight
\(\mathcal W_{ab;cd}(1,2;3,4)\) and write the covariant transition measure as
\begin{equation}
  d\Gamma_{ab;cd}(1;2,3,4)
  =
  d\Pi_b(p_2)d\Pi_c(p_3)d\Pi_d(p_4)\,
  (2\pi)^D\delta^{(D)}(p_1+p_2-p_3-p_4)\,
  \mathcal W_{ab;cd}(1,2;3,4).
  \label{eq:kinetic-transition-measure}
\end{equation}
The weight includes the shell-projection factor appropriate to the covariant
equation \(p^\mu\partial_\mu f=\mathcal C[f]\), spin and color sums or
averages, identical-particle divisors, the ordered-versus-quotient convention
for repeated final states, and the regulator or medium corrections that are
kept at the kinetic scale.  In the scalar full-product convention derived
above,
\[
  \mathcal W_{\phi\phi;\phi\phi}=\frac{\lambda_R^2}{4}.
\]
The channelwise detailed-balance and H-theorem statements below assume the
microreversibility condition
\begin{equation}
  \mathcal W_{ab;cd}(1,2;3,4)
  =
  \mathcal W_{cd;ab}(3,4;1,2)
  \label{eq:kinetic-transition-microreversibility}
\end{equation}
whenever both sides describe the same physical channel.  If a time-reversal
violating theory or a coarse graining produces distinct forward and reverse
weights, the collision operator must be written as a network of separately
weighted gain and loss terms; the pairwise proof below is then replaced by a
KMS/unitarity statement about the full network.
With this convention, the \(2\to2\) contribution to \(\mathcal C_a\) is
\begin{align}
  \mathcal C_a(p_1)
  &=
  \sum_{bcd}\int d\Gamma_{ab;cd}(1;2,3,4)
  \nonumber\\
  &\quad\times
  \Big[
  f_c(p_3)f_d(p_4)
  (1+\eta_a f_a(p_1))(1+\eta_b f_b(p_2))
  \nonumber\\
  &\qquad\qquad
  -
  f_a(p_1)f_b(p_2)
  (1+\eta_c f_c(p_3))(1+\eta_d f_d(p_4))
  \Big].
  \label{eq:kinetic-2to2-collision}
\end{align}
The sum over \(b,c,d\) is a sum over physical channels in the chosen measure
convention.  If the initial or final pair is represented by a quotient rather
than by the ordered product measure, the corresponding finite divisor is
already included in \(\mathcal W\) or in the declared transition measure.
The entropy symmetrization below introduces no further channel-counting
divisor.  When the weight is expressed in terms of an on-shell scattering
amplitude, the amplitude is in the normalization of the scattering chapters
and all finite symmetry and shell-projection factors remain part of
\(\mathcal W\).  In gauge theories the regulator and medium clause is
essential because soft exchange and collinear splitting can contribute at the
same parametric order as hard \(2\to2\) scattering.

\section{Detailed Balance}

The local equilibrium distribution is
\begin{equation}
  f_a^{(0)}(X,\mathbf p)
  =
  \frac{1}{\exp[\beta(X)(-u_\mu(X)p^\mu-\mu_A(X)q^A_a)]-\eta_a}.
  \label{eq:kinetic-local-equilibrium-distribution}
\end{equation}
At a fixed point \(X\), the identity
\begin{equation}
  1+\eta_a f_a^{(0)}
  =
  e^{\beta(-u_\mu p^\mu-\mu_Aq^A_a)}f_a^{(0)}
  \label{eq:kinetic-one-plus-eta-f}
\end{equation}
implies detailed balance.  Indeed, for an allowed reaction satisfying
\[
  p_1+p_2=p_3+p_4,
  \qquad
  q_a^A+q_b^A=q_c^A+q_d^A,
\]
the loss product
\[
  X_{12\to34}
  =
  f_a^{(0)}(p_1)f_b^{(0)}(p_2)
  (1+\eta_c f_c^{(0)}(p_3))(1+\eta_d f_d^{(0)}(p_4))
\]
equals the gain product
\[
  Y_{34\to12}
  =
  f_c^{(0)}(p_3)f_d^{(0)}(p_4)
  (1+\eta_a f_a^{(0)}(p_1))(1+\eta_b f_b^{(0)}(p_2)).
\]
This is an algebraic consequence of
\eqref{eq:kinetic-one-plus-eta-f}: the exponential multiplying
\(f_1f_2f_3f_4\) is the same on the two sides by energy-momentum and charge
conservation.  Therefore the collision term vanishes pointwise for the local
equilibrium form at each fixed \(X\) for the channelwise reversible kernel
\eqref{eq:kinetic-2to2-collision}.  Gradients do not make the local collision
term nonzero; they make the streaming term \(p^\mu\partial_\mu f^{(0)}\)
nonzero.  First-order hydrodynamics solves the linearized collision equation
with this streaming term as the source.  If gain and loss are assigned
different transition weights, the vanishing statement is no longer a
one-channel algebraic identity; it is a constraint on the full reaction
network.

\section{Kinetic Entropy and the H Theorem}

Define the entropy current
\begin{equation}
  S^\mu_{\rm kin}
  =
  -\sum_a\int 2d\Pi_a\,p^\mu
  \left[
    f_a\log f_a
    -\eta_a(1+\eta_af_a)\log(1+\eta_af_a)
  \right].
  \label{eq:kinetic-entropy-current}
\end{equation}
Using the Boltzmann equation,
\begin{equation}
  \partial_\mu S^\mu_{\rm kin}
  =
  -\sum_a\int 2d\Pi_a\,\mathcal C_a[f]\,
  \log\!\left(\frac{f_a}{1+\eta_af_a}\right).
  \label{eq:kinetic-entropy-divergence-before-sym}
\end{equation}
For the \(2\to2\) kernel with nonnegative, microreversible transition measure
\eqref{eq:kinetic-transition-measure}--\eqref{eq:kinetic-transition-microreversibility},
the overall coefficient follows from the same normalization.  Set
\[
  r_i=\frac{f_i}{1+\eta_i f_i}.
\]
For one physical channel, write
\(X=X_{12\to34}\) and \(Y=Y_{34\to12}\).  Then
\[
  \partial_\mu S^\mu_{\rm kin}
  =
  2\int
  d\Pi_1d\Pi_2d\Pi_3d\Pi_4\,
  (2\pi)^D\delta^{(D)}(\cdots)
  \mathcal W_{12;34}
  (X-Y)\log r_1 ,
\]
where the factor \(2\) is the one in
\eqref{eq:kinetic-entropy-divergence-before-sym}.  Incoming exchange,
outgoing exchange, and \(12\leftrightarrow34\) microreversibility give the
identity under the transition integral
\[
  \int \mathcal W(X-Y)\log r_1
  =
  \frac14\int \mathcal W(X-Y)
  \log\!\left(\frac{r_1r_2}{r_3r_4}\right)
  =
  \frac14\int \mathcal W(X-Y)\log\!\left(\frac XY\right).
\]
Thus the four-leg symmetrization supplies \(1/4\), while the entropy current
has already supplied the factor \(2\).  Symmetrizing over the external legs
therefore gives
\begin{align}
  \partial_\mu S^\mu_{\rm kin}
  &=
  \frac12
  \sum_{abcd}\int
  d\Pi_1d\Pi_2d\Pi_3d\Pi_4\,
  (2\pi)^D\delta^{(D)}(p_1+p_2-p_3-p_4)
  \mathcal W_{ab;cd}(1,2;3,4)
  \nonumber\\
  &\quad\times
  (X_{12\to34}-Y_{34\to12})
  \log\!\left(\frac{X_{12\to34}}{Y_{34\to12}}\right)\ge0.
  \label{eq:kinetic-h-theorem}
\end{align}
For the weak scalar full-product convention
\(\mathcal W_{\phi\phi;\phi\phi}=\lambda_R^2/4\), the coefficient multiplying
the ordered four-particle measure in
\eqref{eq:kinetic-h-theorem} is therefore \(\lambda_R^2/8\).
The inequality follows from
\[
  (x-y)\log(x/y)\ge0
  \qquad
  (x>0,\ y>0).
\]
Equality holds, within the set of reactions included in the kernel, exactly
when each allowed gain product equals the corresponding loss product.  When
the reactions connect the species space sufficiently, this condition forces
\(\log[f_a/(1+\eta_af_a)]\) to be a linear combination of the conserved
collision invariants \(1\), \(p^\mu\), and \(q_a^A\), giving
\eqref{eq:kinetic-local-equilibrium-distribution}.

\section{Hydrodynamic Moments}

The kinetic stress tensor and conserved currents are
\begin{align}
  T^{\mu\nu}(X)
  &=
  \sum_a\int 2d\Pi_a\,p^\mu p^\nu f_a(X,\mathbf p),
  \label{eq:kinetic-stress-tensor}\\
  J^\mu_A(X)
  &=
  \sum_a q_{A,a}\int 2d\Pi_a\,p^\mu f_a(X,\mathbf p).
  \label{eq:kinetic-current}
\end{align}
The same shell Jacobian is being used here as in the Wigner projection:
\(2d\Pi_a\,p_a^0=d^d p/(2\pi)^d\) for the density components.
Multiplying \eqref{eq:kinetic-covariant-boltzmann} by \(2p^\nu\) and
integrating gives
\[
  \partial_\mu T^{\mu\nu}
  =
  \sum_a\int 2d\Pi_a\,p^\nu\mathcal C_a[f].
\]
The right side vanishes because each collision conserves total
four-momentum.  Similarly,
\[
  \partial_\mu J^\mu_A
  =
  \sum_a q_{A,a}\int 2d\Pi_a\,\mathcal C_a[f]=0
\]
when the collision channels conserve charge \(q_A\).  Thus the hydrodynamic
Ward identities are the collision-invariant moments of the kinetic equation.

At local equilibrium \eqref{eq:kinetic-local-equilibrium-distribution}, the
moments have the ideal-fluid form
\[
  T^{\mu\nu}_{(0)}=\varepsilon u^\mu u^\nu+p\Delta^{\mu\nu},
  \qquad
  J^\mu_{A(0)}=n_Au^\mu,
\]
with \(p\), \(\varepsilon\), and \(n_A\) obtained from the phase-space
integrals.  First-order transport arises from the first gradient correction
to \(f_a\).

\section{Linearized Collision Operator}

Write
\begin{equation}
  f_a=f_a^{(0)}+f_a^{(0)}(1+\eta_af_a^{(0)})\chi_a .
  \label{eq:kinetic-linearized-ansatz}
\end{equation}
To first order,
\[
  \mathcal C_a[f]=-\sum_b\mathcal L_{ab}\chi_b+O(\chi^2).
\]
The natural inner product on perturbations is
\begin{equation}
  \langle \phi,\psi\rangle
  =
  \sum_a\int 2d\Pi_a\,
  f_a^{(0)}(1+\eta_af_a^{(0)})\,\phi_a(p)\psi_a(p).
  \label{eq:kinetic-linearized-inner-product}
\end{equation}
For the \(2\to2\) collision kernel,
\begin{align}
  \langle \chi,\mathcal L\chi\rangle
  &=
  \frac14
  \sum_{abcd}\int
  d\Pi_1d\Pi_2d\Pi_3d\Pi_4\,
  (2\pi)^D\delta^{(D)}(p_1+p_2-p_3-p_4)
  |\mathcal M_{ab\to cd}|^2
  \nonumber\\
  &\quad\times
  f_1^{(0)}f_2^{(0)}
  (1+\eta_c f_3^{(0)})(1+\eta_d f_4^{(0)})
  \bigl(\chi_1+\chi_2-\chi_3-\chi_4\bigr)^2\ge0.
  \label{eq:kinetic-linearized-positive-form}
\end{align}
The null space consists of perturbations for which
\(\chi_1+\chi_2-\chi_3-\chi_4=0\) for every included reaction.  Under the
same connectivity assumption used in the entropy argument, this null space
is spanned by conserved collision invariants.  Transport coefficients are
quadratic forms involving the inverse of \(\mathcal L\) on the orthogonal
complement of this null space.

\section{Finite Collision Algebra Behind the Integral Formula}
\label{sec:finite-collision-algebra-behind-kinetic-integral}

The continuum formulas above are compact notations for a positive measure on
reaction kinematics.  The sign, null-space, and detailed-balance statements
are already present at the finite algebraic level, so we record that level
explicitly.  This also fixes the precise content of the phrase ``collision
invariant'' before functional analysis is added.

\begin{definition}[Finite reversible collision datum]
\label{def:finite-reversible-collision-datum}
A finite reversible collision datum consists of:
\begin{enumerate}
\item a finite set \(I\) of one-particle cells, each with a statistics sign
  \(\eta_i\in\{+1,-1\}\);
\item an open state domain
  \[
    \mathcal D
    =
    \{f\in\mathbb R^I:\ f_i>0,\ 1+\eta_i f_i>0\ \text{for all }i\};
  \]
\item a finite set \(\mathcal R\) of reversible reactions
  \(r=(i,j;k,l)\), interpreted as \(i+j\rightleftarrows k+l\), with weights
  \(w_r>0\);
\item additive labels \(P_i\) taking values in a real vector space
  \(\mathcal P\), such that
  \begin{equation}
    P_i+P_j=P_k+P_l
    \label{eq:finite-collision-additive-conservation}
  \end{equation}
  for every reaction \(r=(i,j;k,l)\).
\end{enumerate}
For \(f\in\mathcal D\), define
\begin{equation}
  L_r(f)=f_if_j(1+\eta_k f_k)(1+\eta_l f_l),
  \qquad
  G_r(f)=f_kf_l(1+\eta_i f_i)(1+\eta_j f_j),
  \label{eq:finite-collision-loss-gain-products}
\end{equation}
and the stoichiometric vector
\begin{equation}
  \nu_r=-e_i-e_j+e_k+e_l\in\mathbb R^I .
  \label{eq:finite-collision-stoichiometric-vector}
\end{equation}
The collision vector field is
\begin{equation}
  \dot f
  =
  \sum_{r\in\mathcal R}w_r\bigl(L_r(f)-G_r(f)\bigr)\nu_r .
  \label{eq:finite-collision-vector-field}
\end{equation}
\end{definition}

\begin{proposition}[Finite collision conservation and entropy production]
\label{prop:finite-collision-conservation-h-theorem}
For every finite reversible collision datum, the vector field
\eqref{eq:finite-collision-vector-field} preserves every additive conserved
quantity
\begin{equation}
  Q_A(f)=\sum_{i\in I}A(P_i)f_i,
  \qquad A\in\mathcal P^* .
  \label{eq:finite-collision-additive-charge}
\end{equation}
Moreover the entropy
\begin{equation}
  S(f)
  =
  -\sum_{i\in I}
  \left[
    f_i\log f_i
    -\eta_i(1+\eta_if_i)\log(1+\eta_if_i)
  \right]
  \label{eq:finite-collision-entropy}
\end{equation}
satisfies
\begin{equation}
  \dot S(f)
  =
  \sum_{r\in\mathcal R}
  w_r\bigl(L_r(f)-G_r(f)\bigr)
  \log\!\left(\frac{L_r(f)}{G_r(f)}\right)
  \ge0 .
  \label{eq:finite-collision-entropy-production}
\end{equation}
Equality holds if and only if \(L_r(f)=G_r(f)\) for every reaction with
\(w_r>0\).
\end{proposition}

\begin{proof}
The finite state space lets us work entirely inside the open domain
\(\mathcal D\), where every factor in \(L_r\) and \(G_r\) is positive.  Thus
the logarithms below are honest finite-dimensional functions, and the
boundary of \(\mathcal D\) is not used in the argument.

For the conserved quantity,
\[
  \frac{d}{dt}Q_A(f)
  =
  \sum_{r}w_r(L_r-G_r)
  \bigl[-A(P_i)-A(P_j)+A(P_k)+A(P_l)\bigr],
\]
and the bracket vanishes by
\eqref{eq:finite-collision-additive-conservation}.

For the entropy, differentiate the one-cell entropy.  The same formula covers
the classical, Bose, and Fermi signs because
\[
  \frac{\dd}{\dd f}
  \left[
    -f\log f+\eta(1+\eta f)\log(1+\eta f)
  \right]
  =
  -\log\!\left(\frac f{1+\eta f}\right),
\]
with the \(\eta=0\) case understood by taking the limit.  Hence
\begin{equation}
  \frac{\partial S}{\partial f_i}
  =
  -\log\!\left(\frac{f_i}{1+\eta_if_i}\right).
  \label{eq:finite-collision-entropy-derivative}
\end{equation}
For \(r=(i,j;k,l)\),
\[
  \nu_r\cdot\nabla S
  =
  \log\!\left(
  \frac{f_if_j(1+\eta_kf_k)(1+\eta_lf_l)}
       {f_kf_l(1+\eta_if_i)(1+\eta_jf_j)}
  \right)
  =
  \log\!\left(\frac{L_r}{G_r}\right).
  \]
Substitution into \(\dot S=\nabla S\cdot\dot f\) gives
\eqref{eq:finite-collision-entropy-production}.  The sign is fixed reaction
by reaction: for \(x,y>0\),
\[
  (x-y)\log(x/y)\ge0,\qquad x>0,\quad y>0 .
  \]
Indeed \(x>y\) implies \(\log x-\log y>0\), while \(x<y\) makes both factors
negative.  Equality in the sum holds exactly when every positive-weight
reaction has \(L_r=G_r\), because all summands are nonnegative.
\end{proof}

\begin{definition}[Finite local equilibrium and collision invariants]
\label{def:finite-local-equilibrium-collision-invariants}
A state \(f^{(0)}\in\mathcal D\) is a finite equilibrium for the datum when
\begin{equation}
  \frac{f_i^{(0)}}{1+\eta_i f_i^{(0)}}
  =
  \exp[-\alpha-A(P_i)]
  \label{eq:finite-collision-equilibrium-form}
\end{equation}
for some \(\alpha\in\mathbb R\) and \(A\in\mathcal P^*\).  A function
\(\chi:I\to\mathbb R\) is a collision invariant when
\begin{equation}
  \chi_i+\chi_j-\chi_k-\chi_l=0
  \label{eq:finite-collision-invariant}
\end{equation}
for every reaction \(i+j\rightleftarrows k+l\).
\end{definition}

\paragraph{Finite linearized collision operator.}
Let \(f^{(0)}\) be a finite equilibrium.  Write
\begin{equation}
  f_i=f_i^{(0)}+f_i^{(0)}(1+\eta_i f_i^{(0)})\chi_i .
  \label{eq:finite-collision-linearization}
\end{equation}
To first order in \(\chi\),
\begin{equation}
  L_r(f)-G_r(f)
  =
  M_r
  \bigl(\chi_i+\chi_j-\chi_k-\chi_l\bigr)+O(\chi^2),
  \label{eq:finite-collision-linearized-rate}
\end{equation}
where
\begin{equation}
  M_r=L_r(f^{(0)})=G_r(f^{(0)})>0 .
  \label{eq:finite-collision-equilibrium-weight}
\end{equation}
Consequently the linearized entropy production is
\begin{equation}
  \dot S^{(2)}
  =
  \sum_{r\in\mathcal R}
  w_rM_r
  \bigl(\chi_i+\chi_j-\chi_k-\chi_l\bigr)^2\ge0.
  \label{eq:finite-collision-linearized-entropy-production}
\end{equation}
Its null space is precisely the space of collision invariants.  If the
reaction graph is connected enough that every collision invariant is of the
form \(\chi_i=\alpha+A(P_i)\), then the only zero modes are the conserved
hydrodynamic variables.

The linearization is finite-dimensional algebra.  At equilibrium, detailed
balance follows from
\eqref{eq:finite-collision-equilibrium-form} and additive conservation:
\[
  \frac{L_r(f^{(0)})}{G_r(f^{(0)})}
  =
  \exp[-A(P_i)-A(P_j)+A(P_k)+A(P_l)]
  =
  1 .
  \]
For a variation \(\delta f_i=f_i^{(0)}(1+\eta_i f_i^{(0)})\chi_i\),
\[
  \delta\log\!\left(\frac{f_i}{1+\eta_i f_i}\right)=\chi_i .
  \]
Therefore
\[
  \delta\log(L_r/G_r)
  =
  \chi_i+\chi_j-\chi_k-\chi_l .
  \]
Since \(L_r=G_r=M_r\) at equilibrium, this is equivalent to
\eqref{eq:finite-collision-linearized-rate}: if \(R_r=L_r/G_r\), then
\[
  \delta(L_r-G_r)
  =
  M_r\,\delta\log R_r .
\]
The second variation of the entropy-production summand is obtained from
\[
  (L_r-G_r)\log(L_r/G_r)
  =
  M_r(\delta\log R_r)^2+O(\chi^3),
\]
because both \(L_r-G_r\) and \(\log(L_r/G_r)\) vanish to first order at
equilibrium with the same linear coefficient \(M_r^{-1}\delta(L_r-G_r)\).
Summing the reactions gives
\eqref{eq:finite-collision-linearized-entropy-production}.  Its null space
is the intersection of the kernels of the linear forms
\(\chi_i+\chi_j-\chi_k-\chi_l\), which is precisely the space of collision
invariants by Definition~\ref{def:finite-local-equilibrium-collision-invariants}.
Under the stated connectivity hypothesis this space is the span of the
constant invariant and the additive conserved quantities \(A(P_i)\).

The continuum \(2\to2\) formulas are obtained from
Definition~\ref{def:finite-reversible-collision-datum} by replacing the
finite reaction sum by the invariant positive measure
\[
  d\Pi_1d\Pi_2d\Pi_3d\Pi_4\,
  (2\pi)^D\delta^{(D)}(p_1+p_2-p_3-p_4)
  |\mathcal M_{ab\to cd}|^2 .
\]
The finite theorem is not a derivation of this measure from QFT; it is the
exact algebraic part of the kinetic construction once the collision measure
has been derived or declared.

\section{Worked Example: Relaxation-Time Shear Viscosity}

\begin{controlledapproximation}[Constant relaxation time]
\label{ca:kinetic-constant-relaxation-time}
Consider a single massless classical species in \(d=3\), with
\(f^{(0)}=e^{-p/T}\), no conserved charge, and a collision operator replaced
on the shear channel by
\[
  C^{(t)}_{\rm RTA}[\delta f]=-\frac{\delta f}{\tau_R},
\]
where \(\tau_R\) is a constant.  This replacement preserves the shear
tensor structure and gives a closed worked example.  It is a model for the
positive inverse collision operator on the shear subspace, not a derivation
of a microscopic collision kernel.
\end{controlledapproximation}

Let
\[
  \delta f
  =
  -\tau_R
  f^{(0)}
  \frac{p^ip^j}{2TE}
  \sigma_{ij},
  \qquad
  \sigma_{ij}
  =
  \partial_i v_j+\partial_j v_i
  -\frac23\delta_{ij}\partial_kv_k .
\]
Substitution in the kinetic stress tensor gives
\[
  \delta T^{ij}
  =
  -\eta\,\sigma^{ij},
\]
where angular averaging in three dimensions uses
\[
  \int d\Omega\,\hat p_i\hat p_j\hat p_k\hat p_l
  =
  \frac{4\pi}{15}
  (\delta_{ij}\delta_{kl}+\delta_{ik}\delta_{jl}
  +\delta_{il}\delta_{jk}).
\]
Since \(\sigma^i{}_i=0\), the coefficient is
\begin{equation}
  \eta_{\rm RTA}
  =
  \frac{\tau_R}{15T}
  \int\frac{d^3\mathbf p}{(2\pi)^3}
  p^2e^{-p/T}
  =
  \frac{4}{5}p_{\rm therm}\tau_R,
  \label{eq:kinetic-rta-shear-viscosity}
\end{equation}
with
\[
  p_{\rm therm}
  =
  \frac13\int\frac{d^3\mathbf p}{(2\pi)^3}p\,e^{-p/T}
  =
  \frac{T^4}{\pi^2}.
\]
This example displays the general structure of kinetic transport: a source
term from gradients, inversion of the collision operator on a nonconserved
subspace, and a positive quadratic form for the coefficient.

\section{Gauge Theories and Soft Scales}

In weakly coupled gauge theories, hard particles with momentum of order \(T\)
interact through soft gauge exchange of order \(gT\).  Electric fields are
screened by the Debye scale of Chapter~\ref{ch:thermal-gauge-theory-screening};
magnetic fields at scale \(g^2T\) remain nonperturbative in the static
sector.  Kinetic theory for gauge plasmas therefore contains several
coupled effective descriptions:
\[
  \text{hard quasiparticles}
  \quad\leftrightarrow\quad
  \text{soft collective gauge fields}
  \quad\leftrightarrow\quad
  \text{ultrasoft hydrodynamic modes}.
\]
Small-angle scattering, Landau damping, and collinear splitting processes
must be matched in the same source and regulator convention before a
transport coefficient is a well-defined weak-coupling prediction.

\section{Controlled-Limit Requirements}

A kinetic derivation from QFT requires:
\begin{center}
\begin{tabular}{@{}p{0.23\linewidth}p{0.68\linewidth}@{}}
\textbf{spectral input} &
  narrow quasiparticle peaks with a controlled off-shell remainder,\\[0.25em]
\textbf{scale separation} &
  $\tau_{\rm qp}\ll\tau_{\rm coll}$, $\tau_{\rm mem}\ll\tau_X$, with the
  regime-specific relation between $\tau_X$ and $\tau_{\rm coll}$,\\[0.25em]
\textbf{collision input} &
  on-shell amplitudes or cut self-energies with regulator and medium
  corrections,\\[0.25em]
\textbf{closure input} &
  decay or bounding of connected higher correlators and initial
  correlations,\\[0.25em]
\textbf{conservation laws} &
  collision invariants matching QFT Ward identities,\\[0.25em]
\textbf{error estimate} &
  bounds on gradients, finite width, off-shell tails, memory, and closure
  residuals.
\end{tabular}
\end{center}

\begin{openproblem}[QFT derivation of kinetic theory with errors]
\label{op:qft-kinetic-theory-error-derivation}
Derive relativistic kinetic theory from Schwinger--Keldysh QFT with explicit
error bounds, including quasiparticle projection, collision kernels,
screening, collinear processes in gauge theory, and the passage from kinetic
theory to hydrodynamic transport.
\end{openproblem}
