\chapter{Finite-Temperature Path Integrals}
\label{ch:finite-temperature-path-integrals}

\ConventionsEuclidean

This chapter derives the Euclidean thermal path integral from the operator
definition of equilibrium correlation functions.  The derivation is first a
finite-dimensional statement about a trace-class Gibbs operator.  A field
theory path integral on \(S^1_\beta\times\mathbb R^{D-1}\) is obtained only
after a spatial regulator, an operator-domain convention, and a limiting
procedure have been specified.  The circle, spin structure, Matsubara
frequencies, and chemical-potential twists all follow from the trace formula.

\section{Regulated Meaning of the Thermal Path Integral}

The notation
\[
  \int [D\Phi]\,\ee^{-S_E[\Phi]}
\]
has a precise meaning only after one has specified the class of variables
\(\Phi\), the algebra in which they take values, and the limiting operation.
In this chapter it is used in the following restricted sense.  One begins
with a regulated quantum system with Hilbert space \(\Hilb_{\rm reg}\), a
self-adjoint Hamiltonian \(H_{\rm reg}\), and a trace-class Gibbs operator
\(\ee^{-\beta H_{\rm reg}}\).  The thermal path integral is the finite
time-sliced expression obtained by inserting resolutions of the identity in
\[
  \operatorname{Tr}_{\Hilb_{\rm reg}}\ee^{-\beta H_{\rm reg}},
  \label{eq:regulated-thermal-trace-start}
\]
or in trace insertions with operators.  Removing the regulator is a separate
problem.  It may be a lattice thermodynamic limit, a constructive continuum
limit, a Hamiltonian limit, or a perturbative renormalized limit; the symbol
\([D\Phi]\) alone does not specify which one is intended.

There are two algebraically different kinds of variables.  Bosonic variables
are ordinary commuting coordinates or fields at the regulator scale.
Fermionic variables in a coherent-state path integral are generators of a
finite Grassmann algebra at the regulator scale.  Their ``integration'' is a
Berezin linear functional.  It is not a Borel measure on a set of fermionic
field configurations.  The continuum notation for fermionic fields therefore
denotes the limit of a sequence of finite Grassmann-algebra calculations, not
an ordinary measure space of anticommuting points.

\section{Gibbs Traces and Euclidean Correlators}

Let \(\Hilb\) be a Hilbert space and let \(H\) be a self-adjoint Hamiltonian
bounded below.  Fix \(\beta>0\) and assume, in this finite-volume discussion,
that
\[
  Z(\beta)=\operatorname{Tr}_{\Hilb}\ee^{-\beta H}<\infty .
\]
The Gibbs state is the positive normalized functional
\[
  \omega_\beta(A)
  =
  Z(\beta)^{-1}\operatorname{Tr}_{\Hilb}(\ee^{-\beta H}A)
\]
on bounded operators \(A\), and on suitable unbounded operators after a
common trace-class domain has been declared.  For Heisenberg operators
\[
  A(t)=\ee^{\ii Ht}A\ee^{-\ii Ht},
\]
their Euclidean-time translates are
\[
  A_E(\tau)=\ee^{\tau H}A\ee^{-\tau H},
  \qquad 0\leq \tau\leq\beta,
\]
whenever the displayed expression is defined as a quadratic form or as a
trace insertion.  If
\[
  0<\tau_1<\tau_2<\cdots<\tau_n<\beta ,
\]
the Euclidean thermal correlator is
\begin{equation}
  G_E(\tau_1,\ldots,\tau_n)
  =
  Z(\beta)^{-1}
  \operatorname{Tr}\!\left(
    \ee^{-\beta H}
    A_{n,E}(\tau_n)\cdots A_{1,E}(\tau_1)
  \right).
  \label{eq:thermal-euclidean-correlator-trace}
\end{equation}
Moving the factor \(\ee^{-\beta H}\) cyclically through the trace shifts one
operator by \(\beta\) in imaginary time.  This is the finite-volume operator
origin of the KMS boundary condition in
Chapter~\ref{ch:kms-states-thermal-correlators}.

\begin{proposition}[Trace cyclicity gives the thermal circle]
\label{prop:trace-cyclicity-thermal-circle}
Let \(A\) and \(B\) be bounded operators such that the traces below exist.
For \(0<\tau<\beta\),
\[
  \omega_\beta(A_E(\tau)B)
  =
  \omega_\beta(B A_E(\tau-\beta)).
\]
Equivalently, after analytic continuation \(t=-\ii\tau\), the two boundary
values of the strip correlator differ by a cyclic permutation of the
operators.
\end{proposition}

\begin{proof}
Using \(A_E(\tau)=\ee^{\tau H}A\ee^{-\tau H}\),
\[
\begin{aligned}
  \operatorname{Tr}(\ee^{-\beta H}A_E(\tau)B)
  &=
  \operatorname{Tr}(\ee^{-(\beta-\tau)H}A\ee^{-\tau H}B)\\
  &=
  \operatorname{Tr}(B\ee^{-(\beta-\tau)H}A\ee^{-\tau H})\\
  &=
  \operatorname{Tr}(\ee^{-\beta H}B\,\ee^{(\tau-\beta)H}
  A\ee^{-(\tau-\beta)H}).
\end{aligned}
\]
Dividing by \(Z(\beta)\) gives the stated identity.
\end{proof}

\section{Time Slicing and Bosonic Boundary Conditions}

Let \(q=(q^1,\ldots,q^r)\) be a configuration coordinate for a regulated
bosonic quantum system.  Let \(\{|q\rangle\}\) be a generalized coordinate
resolution of the identity.  For
\[
  \beta=N\epsilon ,
\]
the partition function is
\begin{equation}
  Z(\beta)
  =
  \int \prod_{j=1}^{N}dq_j\,
  \prod_{j=1}^{N}
  K_\epsilon(q_{j+1},q_j),
  \qquad
  q_{N+1}=q_1,
  \label{eq:thermal-bosonic-time-slicing}
\end{equation}
where
\[
  K_\epsilon(q',q)=\langle q'|\ee^{-\epsilon H}|q\rangle .
\]
The equality is the trace written as
\[
  \int dq_1\,\langle q_1|(\ee^{-\epsilon H})^N|q_1\rangle
\]
with \(N-1\) intermediate identity insertions.  The condition
\(q_{N+1}=q_1\) is therefore exact at finite \(N\).

For a local bosonic field with a spatial regulator, the same formula has one
coordinate \(q_j\) for every spatial lattice degree of freedom at Euclidean
time slice \(j\).  A formal continuum notation for the result is
\[
  Z(\beta)
  =
  \int_{\phi(\tau+\beta,\mathbf x)=\phi(\tau,\mathbf x)}
  [D\phi]_{\mathrm{reg}}\,
  \ee^{-S_E[\phi]} .
\]
The subscript \(\mathrm{reg}\) is part of the formula: it denotes the
particular lattice, constructive, Hamiltonian, or perturbative regulator from
which the expression is obtained.  The Euclidean time direction is a circle
because the trace identifies the final and initial bosonic coordinates.

\section{Fermionic Coherent States and Spin Structure}

Let \(c_i,c_i^\dagger\), \(i=1,\ldots,r\), be fermionic annihilation and
creation operators with
\[
  \{c_i,c_j^\dagger\}=\delta_{ij},
  \qquad
  \{c_i,c_j\}=\{c_i^\dagger,c_j^\dagger\}=0.
\]
Let \(\eta_i,\bar\eta_i\) be independent Grassmann variables.  The coherent
state
\[
  |\eta\rangle=\ee^{-\sum_i \eta_i c_i^\dagger}|0\rangle
\]
satisfies
\[
  c_i|\eta\rangle=\eta_i|\eta\rangle ,
\]
where the sign comes from moving the odd operator \(c_i\) through the odd
coefficient \(\eta_i\).  The resolution of identity has the form
\[
  \mathbf 1
  =
  \int \prod_i d\bar\eta_i\,d\eta_i\,
  \ee^{-\bar\eta\eta}\,
  |\eta\rangle\langle\eta| ,
  \qquad
  \bar\eta\eta=\sum_i\bar\eta_i\eta_i .
\]

\begin{definition}[Berezin integral at finite regulator]
\label{def:finite-berezin-integral}
Let \(\Lambda\) be the Grassmann algebra generated by
\(\eta_1,\ldots,\eta_r,\bar\eta_1,\ldots,\bar\eta_r\).  The ordered Berezin
integral
\[
  \int \prod_{i=1}^r d\bar\eta_i\,d\eta_i:\Lambda\longrightarrow\mathbb C
\]
is the linear functional fixed by
\[
  \int \prod_{i=1}^r d\bar\eta_i\,d\eta_i\,
  \prod_{i=1}^r \bar\eta_i\eta_i
  =
  (-1)^r
  \label{eq:finite-berezin-normalization}
\]
and by the requirement that every monomial missing at least one generator
integrates to zero.  Equivalently,
\[
  \int \prod_{i=1}^r d\bar\eta_i\,d\eta_i\,
  \exp\!\left(-\sum_i\bar\eta_i\eta_i\right)
  =
  1 .
  \label{eq:finite-berezin-gaussian-normalization}
\]
The variables \(\eta_i,\bar\eta_i\) are algebra generators, not complex
numbers.  The integral is a linear functional on \(\Lambda\), not a positive
measure.
\end{definition}

\begin{lemma}[Fermionic trace identity]
\label{lem:fermionic-coherent-state-trace}
For every operator \(O\) on the finite fermionic Fock space,
\[
  \operatorname{Tr}O
  =
  \int d\bar\eta\,d\eta\,
  \ee^{-\bar\eta\eta}\,
  \langle-\eta|O|\eta\rangle ,
\]
whereas
\[
  \operatorname{Tr}((-1)^F O)
  =
  \int d\bar\eta\,d\eta\,
  \ee^{-\bar\eta\eta}\,
  \langle\eta|O|\eta\rangle .
\]
\end{lemma}

\begin{proof}
It is enough to check one fermionic mode, since the \(r\)-mode formula is the
ordered tensor product of one-mode identities with the ordering in
Definition~\ref{def:finite-berezin-integral}.  In the basis
\(|0\rangle,|1\rangle=c^\dagger|0\rangle\), expand \(O\) as a \(2\times2\)
matrix and use
\[
  |\eta\rangle=|0\rangle-\eta|1\rangle,
  \qquad
  \langle\eta|=\langle0|-\langle1|\bar\eta .
\]
Write \(O_{ab}=\langle a|O|b\rangle\).  Since
\(\langle-\eta|=\langle0|+\langle1|\bar\eta\),
\[
  \ee^{-\bar\eta\eta}\langle-\eta|O|\eta\rangle
  =
  O_{00}-O_{01}\eta+\bar\eta O_{10}
  -(O_{00}+O_{11})\bar\eta\eta .
\]
With the convention
\(\int d\bar\eta\,d\eta\,\bar\eta\eta=-1\), the integral is
\(O_{00}+O_{11}=\operatorname{Tr}O\).  Similarly,
\[
  \ee^{-\bar\eta\eta}\langle\eta|O|\eta\rangle
  =
  O_{00}-O_{01}\eta-\bar\eta O_{10}
  -(O_{00}-O_{11})\bar\eta\eta ,
\]
whose integral is \(O_{00}-O_{11}=\operatorname{Tr}((-1)^F O)\).
\end{proof}

Applying Lemma~\ref{lem:fermionic-coherent-state-trace} to
\((\ee^{-\epsilon H})^N\) gives the thermal Grassmann time-sliced integral.
The first and last coherent-state labels obey
\[
  \eta_{N+1}=-\eta_1,\qquad
  \bar\eta_{N+1}=-\bar\eta_1
\]
for the ordinary thermal partition function.  Thus fermionic variables in
the thermal trace are antiperiodic on \(S^1_\beta\):
\[
  \psi(\tau+\beta,\mathbf x)=-\psi(\tau,\mathbf x),
  \qquad
  \bar\psi(\tau+\beta,\mathbf x)=-\bar\psi(\tau,\mathbf x).
\]
The graded trace \(\operatorname{Tr}((-1)^F\ee^{-\beta H}\cdots)\) uses the
periodic spin structure.  This is the operator origin of the distinction
between a thermal partition function and a supersymmetric index.

\begin{figure}[H]
\centering
\begin{tikzpicture}[>=Stealth,scale=1.0]
  \draw[thick] (0,0) circle (1.25);
  \draw[->,thick] (1.25,0) arc (0:55:1.25);
  \node at (0,0) {\(S^1_\beta\)};
  \node[below] at (0,-1.65) {trace gluing};
  \node[right] at (1.55,0.35) {\(\tau=0\)};
  \node[left] at (-1.55,0.35) {\(\tau=\beta\)};
  \draw[blue!65!black,thick] (3.2,0.75) -- (5.9,0.75);
  \draw[blue!65!black,thick] (3.2,-0.75) -- (5.9,-0.75);
  \draw[->,thick] (5.9,0.75) arc (90:-90:0.75);
  \draw[->,thick] (3.2,-0.75) arc (270:90:0.75);
  \node[right] at (6.05,0.75) {boson: \(+\)};
  \node[right] at (6.05,-0.75) {thermal fermion: \(-\)};
  \node at (4.55,0) {spin structure};
\end{tikzpicture}
\caption{The thermal trace glues the end of the Euclidean time interval to
its beginning.  Coordinate variables are periodic, while ordinary thermal
fermionic coherent-state variables are antiperiodic because the trace
identity contains \(\langle-\eta|\).}
\label{fig:thermal-trace-spin-structure}
\end{figure}

\section{Matsubara Modes as Circle Eigenfunctions}

Let \(f(\tau)\) be a field component on the Euclidean thermal circle obeying
\[
  f(\tau+\beta)=s f(\tau),\qquad s=\pm1.
\]
An eigenfunction of \(-\ii\partial_\tau\) has the form
\(\ee^{\ii\omega\tau}\).  The boundary condition requires
\[
  \ee^{\ii\omega\beta}=s .
\]
Hence
\[
  \omega_n^{\rm B}=\frac{2\pi n}{\beta},
  \qquad n\in\mathbb Z,
\]
for periodic bosonic variables, and
\[
  \omega_n^{\rm F}=\frac{(2n+1)\pi}{\beta},
  \qquad n\in\mathbb Z,
\]
for antiperiodic fermionic variables.

\begin{proposition}[Free scalar thermal Green function]
\label{prop:free-scalar-thermal-green-function}
Let
\[
  K=-\partial_\tau^2-\Delta_{\mathbf x}+m^2
\]
on periodic functions on \(S^1_\beta\times\mathbb R^{D-1}\).  The translation
invariant inverse of \(K\), as a distribution, is
\[
  G_E(\tau,\mathbf x)
  =
  \frac{1}{\beta}\sum_{n\in\mathbb Z}
  \int \frac{d^{D-1}k}{(2\pi)^{D-1}}\,
  \frac{\ee^{\ii\omega_n^{\rm B}\tau+\ii\mathbf k\cdot\mathbf x}}
       {(\omega_n^{\rm B})^2+\mathbf k^2+m^2}.
\]
\end{proposition}

\begin{proof}
The functions
\[
  \beta^{-1/2}\ee^{\ii\omega_n^{\rm B}\tau}
  \]
form the Fourier basis of \(L^2(S^1_\beta)\), and plane waves form the
Fourier basis in the spatial directions.  Acting with \(K\) on one mode
multiplies it by
\[
  (\omega_n^{\rm B})^2+\mathbf k^2+m^2 .
\]
The displayed kernel is therefore the spectral inverse of \(K\) on each
Fourier mode.  Applying \(K\) to it gives the periodic delta distribution
on \(S^1_\beta\) times \(\delta^{D-1}(\mathbf x)\).
\end{proof}

For fermions, the same diagonalization applies to the chosen Euclidean Dirac
operator, but the frequency set is shifted to
\(\omega_n^{\rm F}\).  The shift is a spin-structure effect, not a
perturbative convention.

\section{Thermal Spectral Representation and Analytic Continuation}

The Euclidean circle values determine real-time response only after the
analytic continuation problem has been specified.  The finite-volume spectral
formula makes the required data explicit.  Let \(A\) and \(B\) be bosonic
operators, and let \(|n\rangle\) be an eigenbasis of \(H\) with
\[
  H|n\rangle=E_n|n\rangle .
\]
For \(0<\tau<\beta\),
\[
  G_E^{AB}(\tau)
  =
  Z(\beta)^{-1}
  \sum_{m,n}
  \ee^{-\beta E_n}
  \ee^{-\tau(E_m-E_n)}
  A_{nm}B_{mn}.
\]
The terms with \(E_m=E_n\) are independent of \(\tau\).  Denote this
zero-frequency contribution by
\[
  G^{AB}_0
  =
  Z(\beta)^{-1}
  \sum_{m,n:E_m=E_n}
  \ee^{-\beta E_n}A_{nm}B_{mn}.
\]
For the remaining transitions define the thermal spectral distribution
\begin{equation}
  \rho_{AB}(\omega)
  =
  Z(\beta)^{-1}
  \sum_{m,n:E_m\neq E_n}
  \bigl(\ee^{-\beta E_n}-\ee^{-\beta E_m}\bigr)
  A_{nm}B_{mn}\,
  \delta(\omega-(E_m-E_n)).
  \label{eq:thermal-spectral-distribution-finite-volume}
\end{equation}
Then
\begin{equation}
  G_E^{AB}(\tau)
  =
  G^{AB}_0+
  \int_{\mathbb R}
  \frac{\ee^{-\omega\tau}}{1-\ee^{-\beta\omega}}\,
  \rho_{AB}(\omega)\,d\omega ,
  \qquad 0<\tau<\beta .
  \label{eq:thermal-euclidean-spectral-representation}
\end{equation}
Indeed, on the support \(\omega=E_m-E_n\),
\[
  \frac{\ee^{-\beta E_n}-\ee^{-\beta E_m}}
       {1-\ee^{-\beta\omega}}
  =
  \ee^{-\beta E_n}.
\]
The explicit \(G^{AB}_0\) term is necessary: it is invisible in the
commutator spectral density because the numerator in
\eqref{eq:thermal-spectral-distribution-finite-volume} vanishes at
\(\omega=0\).

The bosonic Matsubara transform is
\[
  G_E^{AB}(\ii\omega_\ell^{\rm B})
  =
  \int_0^\beta d\tau\,
  \ee^{\ii\omega_\ell^{\rm B}\tau}
  G_E^{AB}(\tau).
\]
Using \(\ee^{\ii\omega_\ell^{\rm B}\beta}=1\) in
\eqref{eq:thermal-euclidean-spectral-representation} gives
\begin{equation}
  G_E^{AB}(\ii\omega_\ell^{\rm B})
  =
  \beta G^{AB}_0\delta_{\ell 0}
  +
  \int_{\mathbb R}
  \frac{\rho_{AB}(\omega)}{\omega-\ii\omega_\ell^{\rm B}}\,d\omega .
  \label{eq:matsubara-spectral-transform}
\end{equation}
For fermionic operators the graded spectral density and the denominator
\(1+\ee^{-\beta\omega}\) replace the bosonic ones; the antiperiodic identity
\(\ee^{\ii\omega_\ell^{\rm F}\beta}=-1\) gives the same Cauchy-transform
form with \(\omega_\ell^{\rm F}\).

The retarded correlator is the boundary value of a function obtained from
the same spectral density,
\[
  G_R^{AB}(z)
  =
  \int_{\mathbb R}
  \frac{\rho_{AB}(\omega)}{\omega-z}\,d\omega,
  \qquad \operatorname{Im}z>0,
\]
under the growth and distributional hypotheses required for the integral.
Thus the rule
\[
  \ii\omega_n\mapsto \omega+\ii0
\]
is a shorthand for a spectral reconstruction problem.  It is justified when
the Euclidean data determine a spectral distribution with the stated
analytic control; the discrete Matsubara values alone do not constitute such
a proof.

\section{Chemical Potentials, Twists, and Holonomies}

Let \(Q\) be the self-adjoint generator of a global \(U(1)\) symmetry and
assume \([H,Q]=0\).  The grand-canonical Gibbs operator is
\[
  \ee^{-\beta(H-\mu Q)} .
\]
If an operator \(\Phi_q\) has charge \(q\),
\[
  [Q,\Phi_q]=q\Phi_q,
\]
then moving \(\ee^{\beta\mu Q}\) through \(\Phi_q\) gives
\[
  \ee^{\beta\mu Q}\Phi_q\ee^{-\beta\mu Q}
  =
  \ee^{\beta\mu q}\Phi_q .
\]
Consequently the Euclidean coordinate representation may be written in two
equivalent ways:
\[
  D_\tau\longmapsto D_\tau-\mu q
\]
with the original thermal boundary condition, or as a twisted boundary
condition
\[
  \Phi_q(\tau+\beta,\mathbf x)
  =
  s\,\ee^{\beta\mu q}\Phi_q(\tau,\mathbf x),
  \qquad s=+1\ \hbox{for bosons},\quad s=-1\ \hbox{for fermions}.
\]
For imaginary chemical potential \(\mu=\ii\nu\), the twist is a unitary
background holonomy
\[
  \exp(\ii\beta\nu q)
\]
around \(S^1_\beta\).  If the \(U(1)\) is compact and the charges are
integers in the chosen normalization, large background gauge transformations
identify
\[
  \nu\sim \nu+\frac{2\pi}{\beta}.
\]
For real \(\mu\), the Euclidean action of a relativistic theory is generally
complex in common field coordinates; this is the operator origin of the
finite-density sign problem discussed later in this volume.

\section{Gauge Fields on the Thermal Circle}

For a gauge theory, the Euclidean thermal path integral integrates over
connections on \(S^1_\beta\times\Sigma\) modulo gauge transformations whose
periodicity is compatible with the chosen bundle and matter boundary
conditions.  The component \(A_\tau\) has a gauge-invariant zero-mode datum:
the thermal holonomy
\[
  \mathcal P(\mathbf x)
  =
  \operatorname{Pexp}
  \left(\int_0^\beta A_\tau(\tau,\mathbf x)\,d\tau\right),
\]
defined up to conjugation.  In a compact lattice regulator this holonomy is
the Polyakov loop variable developed in
Chapter~\ref{ch:lattice-yang-mills-nonperturbative-definition}.  In a
continuum gauge-fixed coordinate chart it should not be discarded as a
pure-gauge variable: a gauge transformation that removes a constant
\(A_\tau\) may fail to be periodic around the thermal circle.

The gauge-theory thermal trace is over the physical Hilbert space, or
equivalently over gauge-invariant states after imposing Gauss constraints.
In the Euclidean path integral this appears as integration over gauge
holonomies rather than fixing a chemical potential for a local gauge charge.
External background holonomies may be fixed for global symmetries.  Dynamical
gauge holonomies are integrated, and their center transformations encode the
thermal one-form symmetry data discussed in the gauge-theory volumes.

\begin{openproblem}[Euclidean thermal reconstruction]
\label{op:euclidean-thermal-reconstruction}
Give constructive hypotheses under which Schwinger functions on
\(S^1_\beta\times\mathbb R^{D-1}\), with the appropriate spin structure and
background holonomies, reconstruct a real-time KMS state with local
observables, spectral functions, and transport correlators.  The hypotheses
must replace the vacuum spectrum condition by precise thermal analyticity and
growth bounds and must specify the operator topology in which the
infinite-volume limit is taken.
\end{openproblem}
