\chapter{QCD Phase Structure, Plasma, and Dense Matter}
\label{ch:qcd-phase-structure-plasma-dense-matter}

\ConventionsMixed

This chapter records the finite-temperature and finite-density phase
questions of QCD in the same convention as the Yang--Mills chapters:
\[
  S_{\rm YM}
  =
  \frac{1}{4g^2}\int\operatorname{tr}(F_{\mu\nu}F^{\mu\nu}),
  \qquad
  \operatorname{tr}_{\square}(T^aT^b)=\delta^{ab}
\]
for \(SU(N_c)\) unless another invariant bilinear form is explicitly stated.
The chapter is deliberately not a catalogue of phase-diagram lore.  Each
order parameter is tied to a symmetry, a limiting procedure, or a specified
source.  Each weak-coupling statement is separated from the nonperturbative
questions that remain after the weak-coupling scale hierarchy has been
identified.

\section{Thermal QCD Data}

Let \(N_f\) Dirac quarks transform in the fundamental representation of
\(SU(N_c)\).  Put the Euclidean theory on
\[
  M_\beta=S^1_\beta\times \mathbb R^3,
  \qquad
  \beta=T^{-1},
\]
with periodic gauge fields and antiperiodic quark fields around
\(S^1_\beta\).  A baryon chemical potential \(\mu_B\) couples to the
conserved baryon number \(B\), normalized so that a quark has baryon charge
\(1/N_c\).  Equivalently, every quark flavor has quark chemical potential
\[
  \mu_q=\frac{\mu_B}{N_c}.
\]
At finite spatial volume \(V\), a regulator \(a\), and quark masses
\(m=(m_1,\ldots,m_{N_f})\), define
\begin{equation}
  Z_{a,V}(T,\mu_B,m)
  =
  \operatorname{Tr}_{\Hilb_{a,V}}
  \exp\{-\beta(H_{a,V}-\mu_B B_{a,V})\},
  \label{eq:qcd-phase-finite-volume-partition}
\end{equation}
when the Hamiltonian regulator is used.  In a Euclidean lattice regulator
the same object is represented by the corresponding gauge integral with the
temporal links weighted by \(e^{\pm a\mu_q}\) for forward and backward quark
propagation.  The pressure is the thermodynamic limit
\begin{equation}
  p(T,\mu_B,m)
  =
  \lim_{a\to0}\lim_{V\to\infty}
  \frac{T}{V}\log Z_{a,V}(T,\mu_B,m),
  \label{eq:qcd-pressure-definition}
\end{equation}
provided the limit exists along the chosen continuum trajectory.

\begin{definition}[QCD phase specification]
\label{def:qcd-phase-specification}
A QCD phase specification at fixed \(N_c,N_f\) consists of:
\begin{enumerate}
  \item the thermodynamic pressure \(p(T,\mu_B,m)\) as a function on the
  domain where the continuum thermal state is defined;
  \item a list of symmetries that are exact at the chosen masses and chemical
  potentials;
  \item order parameters or source derivatives associated to those exact
  symmetries;
  \item the limiting prescription used to detect spontaneous symmetry
  breaking;
  \item a status label for each nonanalytic locus: theorem, lattice
  continuum-extrapolation claim, controlled weak-coupling claim, or
  conjectural continuation.
\end{enumerate}
The word ``phase'' is used in this chapter only through such a specification.
\end{definition}

\paragraph{Physical-output map of the chapter.}
The rest of the chapter uses Definition~\ref{def:qcd-phase-specification} as
a repeated physical-output test, not as introductory bookkeeping.  First,
finite-regulator analyticity and source curvatures identify which pressure
or susceptibility observable can become nonanalytic in the thermodynamic
continuum limit.  Second, center and chiral sections attach candidate order
parameters to exact symmetries and to specified source limits.  Third, the
plasma and transport sections ask for real-time response observables extracted
from pressure derivatives, spectral limits, and controlled hard-thermal-loop
windows, while keeping those data separate from claims about hydrodynamic
poles.  Fourth, the baryon-density sections explain the sign problem,
imaginary-source diagnostics, and high-density effective theory before any
dense phase is named.  Finally, the CFL analysis proceeds from attractive
Fermi-surface channels to gauge-invariant order diagnostics, screening
response, collective fields, and anomaly matching.  Thus a named region in the
QCD phase diagram is accepted only after the corresponding pressure,
symmetry, source, response, and limiting data have been displayed or
explicitly marked as uncontrolled.

\section{Finite-Regulator Analyticity and Thermodynamic Phases}
\label{sec:qcd-phase-finite-regulator-analyticity}

A phase transition is not a property of a single finite lattice integral.  At
finite cutoff and finite volume the partition function is an analytic
function of its sources away from its zeros.  Nonanalyticity can only emerge
from a limiting procedure: thermodynamic volume, continuum cutoff, chiral
mass, or a source-selection limit.  This elementary point is important
because it prevents a phase diagram from being treated as a drawing whose
curves exist before the limiting observables have been defined.

\begin{definition}[Finite lattice thermal gauge setup]
\label{def:qcd-finite-lattice-thermal-gauge-setup}
Fix a finite Euclidean lattice
\[
  \Lambda=(N_\tau,N_s^3,a)
\]
with compact gauge group \(G\), Haar measure \(dU\) on each link, and
antiperiodic temporal boundary condition for quarks.  Let
\(\kappa=(\kappa_\alpha)\) denote finitely many gauge couplings multiplying
bounded continuous class functions \(S_\alpha(U)\), let
\(\zeta=e^{a\mu_q}\) be the quark fugacity, and let
\[
  M_\Lambda(U;m,\zeta)
\]
be the finite fermion matrix in the chosen lattice discretization, with
entries polynomial in the masses \(m_i\) and Laurent-polynomial in \(\zeta\).
For an integer number of unrooted Dirac flavors, the finite-regulator
partition function is
\begin{equation}
  Z_\Lambda(\kappa,m,\zeta)
  =
  \int_{G^{E(\Lambda)}}\prod_{\ell}dU_\ell\,
  \exp\left[-\sum_\alpha \kappa_\alpha S_\alpha(U)\right]
  \prod_{i=1}^{N_f}\det M_{\Lambda,i}(U;m_i,\zeta),
  \label{eq:qcd-finite-lattice-partition-analytic}
\end{equation}
where \(E(\Lambda)\) is the finite link set.  Rooted or otherwise
non-polynomial fermion prescriptions are not included in this setup unless a
separate branch and locality construction has been supplied.
\end{definition}

\paragraph{Finite-regulator analyticity.}
For the setup of Definition~\ref{def:qcd-finite-lattice-thermal-gauge-setup},
\(Z_\Lambda(\kappa,m,\zeta)\) is entire in the gauge couplings and masses
and Laurent-polynomial in the fugacity \(\zeta\).  Consequently the finite
pressure
\[
  p_\Lambda(\kappa,m,\zeta)
  =
  \frac{T}{V_\Lambda}\log Z_\Lambda(\kappa,m,\zeta)
\]
is holomorphic on every simply connected parameter domain that does not meet
the zero locus of \(Z_\Lambda\), after a branch of the logarithm has been
chosen.  This finite-dimensional analytic fact becomes a statement about a
thermodynamic or continuum pressure only after a separate compact-convergence
input.  The lattice has finitely many links, and \(G^{E(\Lambda)}\) is
compact.  For
each fixed gauge field \(U\), the determinant factors are polynomials in the
masses and Laurent polynomials in \(\zeta\).  The exponential of the gauge
action has an absolutely and uniformly convergent power series in
\(\kappa\) on compact subsets of complex coupling space, because every
\(S_\alpha(U)\) is bounded on the compact integration domain.  Termwise
integration is justified by uniform convergence on compact parameter sets.
Thus the integral defines an entire function in \(\kappa\) and \(m\), with
only finitely many positive and negative fugacity powers.  Wherever this
function is nonzero on a simply connected domain, a holomorphic logarithm
exists, and the displayed finite pressure is holomorphic there.

\begin{definition}[Thermodynamic phase transition and Lee--Yang obstruction]
\label{def:qcd-thermodynamic-phase-transition}
Let \(x\) denote a real collection of thermal parameters, masses, and
sources, and let \(p_\Lambda(x)\) be the finite-regulator pressure along a
specified sequence of lattices and continuum trajectories.  A thermodynamic
phase transition at \(x_0\) means that the limiting pressure
\[
  p(x)=\lim_{\Lambda\to\infty}p_\Lambda(x)
\]
exists in a real neighborhood of \(x_0\) but is not real analytic at
\(x_0\), or that the limiting local Gibbs/KMS state is not unique after the
source-selection prescription has been removed.  The finite-regulator
Lee--Yang locus is
\[
  \mathcal Z_\Lambda
  =
  \{z\in\mathcal U_{\mathbb C}\mid Z_\Lambda(z)=0\}.
\]
\end{definition}

\begin{remark}[Zero-free neighborhoods exclude pressure singularities]
\label{rem:qcd-lee-yang-zero-free-analytic-limit}
Let \(\mathcal U_{\mathbb C}\) be a complex neighborhood of a real point
\(x_0\).  Suppose that, for all sufficiently large \(\Lambda\),
\(Z_\Lambda\) has no zeros in \(\mathcal U_{\mathbb C}\), that the logarithm
branches are chosen compatibly, and that \(p_\Lambda\) converges uniformly on
compact subsets of \(\mathcal U_{\mathbb C}\).  Then the limiting pressure
\(p\) is holomorphic on \(\mathcal U_{\mathbb C}\), hence real analytic near
\(x_0\).  Therefore a pressure nonanalyticity at \(x_0\) requires either
Lee--Yang zeros approaching \(x_0\) in every complex neighborhood, or failure
of the stated uniform analytic convergence.
By the finite-regulator analyticity discussion above, each \(p_\Lambda\) is
holomorphic on \(\mathcal U_{\mathbb C}\) after the stated choice of
logarithm.  The uniform compact convergence hypothesis then makes the limit
holomorphic by the Weierstrass theorem for holomorphic functions.
Restricting a holomorphic function to the real slice gives a real-analytic
function.  The contrapositive gives the final statement: if the real pressure
has a nonanalyticity at \(x_0\), the hypotheses just listed cannot all hold on
any complex neighborhood of \(x_0\).  Thus either zeros approach \(x_0\), or
the finite-regulator analytic functions fail to converge in the compact-open
holomorphic topology there.
\end{remark}

\paragraph{Source curvature as an integrated connected correlator.}
Let \(h\) be a real source coupled at finite regulator to an integrated
gauge-invariant local observable
\[
  X_\Lambda=\sum_{x\in\Lambda_s}a^3\,\mathcal O_x
\]
through the Euclidean weight \(\exp(hX_\Lambda)\).  With
\[
  p_\Lambda(h)=\frac{T}{V_\Lambda}\log Z_\Lambda(h),
  \qquad
  \overline{\mathcal O}_\Lambda=V_\Lambda^{-1}X_\Lambda,
\]
one has
\begin{equation}
  \frac{\partial p_\Lambda}{\partial h}
  =
  T\langle\overline{\mathcal O}_\Lambda\rangle_h,
  \qquad
  \frac{\partial^2 p_\Lambda}{\partial h^2}
  =
  T\,V_\Lambda
  \left(
    \langle\overline{\mathcal O}_\Lambda^2\rangle_h
    -
    \langle\overline{\mathcal O}_\Lambda\rangle_h^2
  \right).
  \label{eq:qcd-source-curvature-susceptibility}
\end{equation}
Thus a divergent susceptibility is a statement about the scaling of an
integrated connected correlator in the thermodynamic limit.

At finite regulator \(X_\Lambda\) is an ordinary measurable function on the
finite configuration space or finite lattice field space, and \(Z_\Lambda(h)\)
is a finite-dimensional integral with positive radius of analyticity in \(h\)
around any real value for which the weight is integrable.  Differentiating
under the integral therefore gives
\[
  \partial_h\log Z_\Lambda(h)=\langle X_\Lambda\rangle_h .
\]
For the second derivative, write
\(\langle Y\rangle_h=Z_\Lambda(h)^{-1}\int Y\,e^{hX_\Lambda}d\mu_\Lambda\).
Then
\[
  \partial_h\langle Y\rangle_h
  =
  \langle YX_\Lambda\rangle_h
  -
  \langle Y\rangle_h\langle X_\Lambda\rangle_h .
\]
Taking \(Y=X_\Lambda\) yields the connected two-point formula
\[
  \partial_h^2\log Z_\Lambda(h)
  =
  \langle X_\Lambda^2\rangle_h-\langle X_\Lambda\rangle_h^2 .
\]
Multiplying by \(T/V_\Lambda\) and writing
\(X_\Lambda=V_\Lambda\overline{\mathcal O}_\Lambda\) gives
\eqref{eq:qcd-source-curvature-susceptibility}.  Thus the possible
thermodynamic singularity is not a pointwise singularity of
\(\mathcal O_x\); it is the volume growth of the integrated connected
correlator after the infrared limit is taken.

For physical QCD with fundamental quarks of nonzero mass, the most familiar
finite-temperature change near \(T\sim\Lambda_{\rm QCD}\) is not attached to
an exact fundamental center symmetry and not attached to an exact chiral
symmetry.  It is therefore a crossover unless a nonanalyticity is established
in the thermodynamic continuum theory.  Pure Yang--Mills theory and chiral
limits have sharper order parameters, and those are treated first.

\section{Center Deconfinement in Pure Yang--Mills}

Let \(G=SU(N_c)\) and first remove all dynamical fundamental matter.  The
thermal Wilson line at spatial point \(\mathbf x\) is
\begin{equation}
  \mathcal P(\mathbf x)
  =
  \mathcal P_\tau
  \exp\left(\ii\int_0^\beta A_0(\tau,\mathbf x)\,d\tau\right).
  \label{eq:qcd-phase-polyakov-holonomy}
\end{equation}
For an irreducible representation \(R\) define the normalized Polyakov loop
\begin{equation}
  P_R(\mathbf x)=\frac{1}{d_R}\operatorname{tr}_R\mathcal P(\mathbf x).
  \label{eq:qcd-phase-polyakov-loop}
\end{equation}
If \(R\) has \(N\)-ality \(k_R\), then the thermal center transformation by
\(z\in Z(SU(N_c))\) sends
\[
  P_R(\mathbf x)\longmapsto z^{k_R}P_R(\mathbf x).
\]
The lattice construction of the transformation and the static-source
interpretation were given in
Chapter~\ref{ch:lattice-yang-mills-nonperturbative-definition}; here we use
them as the phase-ordering language of the thermal theory.

\begin{definition}[Center-order deconfinement]
\label{def:qcd-center-order-deconfinement}
In pure \(SU(N_c)\) Yang--Mills, choose a small source \(h\) that selects a
center sector and set
\[
  \overline P_\square
  =
  V^{-1}\int_{\mathbb R^3}P_\square(\mathbf x)\,d^3x
\]
in a finite box before the infinite-volume limit.  The center order
parameter is
\begin{equation}
  \ell(T)
  =
  \lim_{h\downarrow0}\lim_{V\to\infty}
  \left\langle \overline P_\square\right\rangle_{T,h,V}.
  \label{eq:qcd-phase-polyakov-order}
\end{equation}
The thermal state is center symmetric when \(\ell(T)=0\) and center broken
when \(\ell(T)\ne0\).  A center deconfinement transition is a nonanalytic
change of the infinite-volume free energy across which this symmetry
realization changes.
\end{definition}

\begin{remark}[Center symmetry forces the finite-volume one-point loop]
\label{rem:qcd-center-finite-volume-loop-zero}
At finite spatial volume and zero center-breaking source in pure
\(SU(N_c)\) Yang--Mills,
\[
  \left\langle P_R(\mathbf x)\right\rangle=0
\]
for every representation \(R\) whose \(N\)-ality is nonzero.

The finite-volume regulated measure and action are invariant under the
thermal center transformation.  Averaging the expectation value over the
finite group \(Z(SU(N_c))\) gives
\[
  \langle P_R\rangle
  =
  \frac1{N_c}\sum_{j=0}^{N_c-1}
  e^{2\pi\ii j k_R/N_c}\langle P_R\rangle .
\]
If \(k_R\not\equiv0\pmod{N_c}\), the finite geometric sum vanishes.  Hence
\(\langle P_R\rangle=0\).  A nonzero Polyakov-loop order parameter is
therefore necessarily a spontaneous-symmetry-breaking limit, not a
finite-volume one-point function.
\end{remark}

After multiplicative line renormalization, the one-point function has the
static-source interpretation
\begin{equation}
  \langle P_R\rangle_{\rm ren}
  =
  \exp\{-\beta(F_R-F_0)\}.
  \label{eq:qcd-phase-static-source-free-energy}
\end{equation}
Thus \(\ell=0\) is equivalent to an infinite free-energy cost for an isolated
fundamental external charge, in the center-symmetric limiting state.  This
statement is logically distinct from the zero-temperature Wilson-loop area
law.  The latter diagnoses the large-distance static potential of a
source-antisource pair; the former diagnoses the realization of thermal
center symmetry and the free energy of one wrapped external line.

\begin{remark}[Dynamical fundamental quarks]
\label{rem:qcd-dynamical-quarks-center}
Fundamental quarks explicitly break the fundamental thermal center symmetry.
Consequently \(P_\square\) remains a useful renormalized static-source
observable, but not a strict order parameter.  In physical QCD the rapid
change of the Polyakov loop is evidence for liberation of color degrees of
freedom in a crossover regime.  Existence of an exact phase transition
requires a separate thermodynamic-limit statement.
\end{remark}

\subsection{Polyakov-Loop Effective Measures at Finite Regulator}
\label{sec:qcd-polyakov-loop-effective-measures}

A Polyakov-loop effective theory is not, by definition, a new continuum QCD.
At finite regulator it is a marginal measure obtained from the thermal
lattice gauge theory by retaining only temporal Wilson lines.  Its usefulness
depends on whether the resulting nonlocal measure can be approximated by a
controlled local or rapidly decaying interaction.

\begin{definition}[Exact finite-regulator Polyakov-loop effective measure]
\label{def:qcd-polyakov-loop-effective-measure}
Fix a finite thermal lattice
\(\Lambda=N_\tau\times\Lambda_s\) with spacing \(a\), compact gauge group
\(SU(N_c)\), and thermal circle length \(\beta=aN_\tau\).  For a gauge
configuration \(U\), define the temporal Wilson line at a spatial site
\(\mathbf x\) by
\[
  \mathcal P_{\mathbf x}(U)
  =
  \prod_{\tau=0}^{N_\tau-1}U_0(\tau,\mathbf x),
\]
with the product ordered along the Euclidean time circle.  For every bounded
class function \(F\) of the collection \(\{\mathcal P_{\mathbf x}\}\), define
\[
  \int F(P)\,d\nu_{\Lambda}(P)
  =
  \frac{1}{Z_{\Lambda}}
  \int d\mu_{\Lambda}(U)\,
  F(\{\mathcal P_{\mathbf x}(U)\})\,
  e^{-S_g[U]}\prod_f\det D_f[U].
\]
When the regulated weight is positive, this is a probability measure on
\(SU(N_c)^{\Lambda_s}\), restricted to class functions at each site.  At
real chemical potential, or in any other situation where the determinant is
complex, the same formula defines a normalized complex linear functional
rather than a probability measure.  In either case it is the exact
finite-regulator Polyakov-loop effective object.  If it is absolutely
continuous with
respect to the product Haar measure \(dP=\prod_{\mathbf x}dP_{\mathbf x}\),
one may write
\[
  d\nu_{\Lambda}(P)=Z_{\rm eff}^{-1}e^{-S_{\rm eff,\Lambda}(P)}dP,
\]
where \(S_{\rm eff,\Lambda}\) is defined up to an additive constant and up
to changes on Haar-null sets.
\end{definition}

\paragraph{Character coordinates and center charge.}
Assume \(d\nu_{\Lambda}\) has an \(L^2\) density with respect to product
Haar measure.  Then the density admits an \(L^2\) character expansion
\[
  \frac{d\nu_{\Lambda}}{dP}
  =
  \sum_{\{\lambda_{\mathbf x}\}}
  c_{\{\lambda_{\mathbf x}\}}
  \prod_{\mathbf x\in\Lambda_s}\chi_{\lambda_{\mathbf x}}(P_{\mathbf x}),
\]
where only finitely many sites carry nontrivial representations in each
basis vector.  In pure Yang--Mills, the coefficient
\(c_{\{\lambda_{\mathbf x}\}}\) vanishes unless the total \(N\)-ality
\[
  \sum_{\mathbf x} k_{\lambda_{\mathbf x}}
  \]
is \(0\) modulo \(N_c\).  With dynamical fundamental quarks this selection
rule is explicitly broken by terms of nonzero total \(N\)-ality.

The Peter--Weyl theorem gives an orthonormal basis of class functions on
\(SU(N_c)\) consisting of irreducible characters.  Tensoring these bases
over the finite set \(\Lambda_s\) gives the displayed expansion in
\(L^2(SU(N_c)^{\Lambda_s})\).

For pure Yang--Mills the finite-regulator measure is invariant under the
thermal center transformation
\[
  P_{\mathbf x}\mapsto zP_{\mathbf x},\qquad z\in Z(SU(N_c)),
\]
applied simultaneously at all spatial sites.  Since
\(\chi_{\lambda}(zP)=z^{k_\lambda}\chi_\lambda(P)\), invariance of the
density forces
\[
  c_{\{\lambda_{\mathbf x}\}}
  =
  z^{\sum_{\mathbf x}k_{\lambda_{\mathbf x}}}
  c_{\{\lambda_{\mathbf x}\}}
  \quad\text{for all }z\in Z(SU(N_c)).
\]
This is possible only when the total \(N\)-ality vanishes modulo \(N_c\), or
when the coefficient is zero.  Fundamental quark determinants contain
closed fermion paths winding once around the thermal circle; such paths
carry fundamental \(N\)-ality and therefore break the selection rule.

\begin{controlledapproximation}[Local Polyakov-loop action]
\label{ca:qcd-local-polyakov-loop-action}
A local Polyakov-loop action is a truncation of the exact effective measure
to finitely many character monomials, usually organized by spatial range,
representation size, winding number, and powers of a strong-coupling or
hopping parameter.  The truncation is a controlled approximation only in a
domain where the omitted character coefficients are bounded by a convergent
cluster or hopping expansion, or by a separately stated numerical error
estimate.
\end{controlledapproximation}

\paragraph{Leading strong-coupling tube in pure Yang--Mills.}
Let the pure-gauge lattice action have a plaquette character expansion whose
fundamental coefficient is \(u_{\square}\), normalized so that
\(u_{\square}=0\) at zero plaquette coupling.  In a finite lattice
strong-coupling expansion, the leading center-invariant nearest-neighbor
interaction between Polyakov loops has the form
\[
  S_{\rm eff,\Lambda}^{(1)}(P)
  =
  -J_1
  \sum_{\langle\mathbf x\mathbf y\rangle}
  \left[
    \chi_{\square}(P_{\mathbf x})
    \chi_{\overline\square}(P_{\mathbf y})
    +
    \chi_{\overline\square}(P_{\mathbf x})
    \chi_{\square}(P_{\mathbf y})
  \right],
\]
with
\[
  J_1=u_{\square}^{N_\tau}+O(u_{\square}^{N_\tau+2})
\]
in the same expansion.  The displayed term is center neutral and therefore
compatible with the character-coordinate center-selection rule above.

Expand each plaquette Boltzmann factor in irreducible characters.  A
nonzero contribution to the Polyakov-loop marginal measure must leave no
unpaired spatial link after Haar integration.  For two neighboring spatial
sites \(\mathbf x,\mathbf y\), the minimal surface that connects their
temporal Wilson lines is a tube of \(N_\tau\) temporal-spatial plaquettes
wrapping once around the thermal circle.  Placing the fundamental character
on each plaquette of this tube and integrating the shared spatial links
contracts the representation matrices into
\[
  \chi_{\square}(P_{\mathbf x})
  \chi_{\overline\square}(P_{\mathbf y})
\]
or its complex conjugate.  The weight of the minimal tube is
\(u_{\square}^{N_\tau}\).  Corrections either decorate the tube or use
larger representations and therefore carry higher powers of the plaquette
character coefficients; in the fundamental channel the first plaquette
decorations occur at the indicated higher order.  The product has total
\(N\)-ality \(1-1=0\), so it is invariant under the simultaneous center
transformation.

\paragraph{Leading heavy-quark center-breaking term.}
For Wilson fermions with hopping parameter \(\kappa\) and quark chemical
potential \(\mu_q\), the leading temporal-winding term in the hopping
expansion of \(N_f\) degenerate fundamental Dirac quarks is
\[
  S_{\rm eff,\Lambda}^{\rm hq}(P)
  =
  -h
  \sum_{\mathbf x}
  \left[
    e^{\beta\mu_q}\chi_{\square}(P_{\mathbf x})
    +
    e^{-\beta\mu_q}\chi_{\overline\square}(P_{\mathbf x})
  \right]
  +\cdots,
\]
where, in the standard Wilson hopping normalization,
\[
  h=2N_f(2\kappa)^{N_\tau}.
\]
The ellipsis denotes longer closed fermion paths, multiple windings, spatial
hops, and higher powers of \(\kappa\).  The displayed term has nonzero
\(N\)-ality and is the finite-regulator origin of explicit center breaking
by dynamical fundamental quarks.

Write the Wilson Dirac operator as \(D=1-H\), where \(H\) is the hopping
matrix.  At finite lattice volume,
\[
  \log\det D
  =
  \operatorname{Tr}\log(1-H)
  =
  -\sum_{n\ge1}\frac1n\operatorname{Tr}H^n
\]
as a formal hopping expansion.  Each trace is a sum over closed lattice
paths.  A path that winds once forward around the Euclidean time circle at
fixed \(\mathbf x\) has \(N_\tau\) forward temporal hops and gauge factor
\(\operatorname{tr}_{\square}P_{\mathbf x}\).  The spin projectors along a
straight temporal Wilson loop leave two spin states, giving the factor \(2\);
the \(N_f\) degenerate flavors multiply the result by \(N_f\).  The
antiperiodic fermion boundary condition supplies an additional minus sign
for a path with odd thermal winding.  This sign cancels the minus sign in
the logarithmic expansion, so the leading forward-winding contribution to
\(\log\det D\) is
\[
  2N_f(2\kappa)^{N_\tau}e^{\beta\mu_q}
  \chi_{\square}(P_{\mathbf x}).
\]
Here \(\beta=aN_\tau\), so the \(N_\tau\) factors of \(e^{a\mu_q}\) combine
to \(e^{\beta\mu_q}\).  The backward winding gives the conjugate term with
\(e^{-\beta\mu_q}\) and \(\chi_{\overline\square}\).  Since
\(e^{-S_{\rm eff}}\) contains \(\det D=\exp(\log\det D)\), the same positive
term in \(\log\det D\) appears with a minus sign in \(S_{\rm eff}\), as
displayed.

Under \(P_{\mathbf x}\mapsto zP_{\mathbf x}\), the first term is multiplied
by \(z\) and the second by \(z^{-1}\).  Hence the heavy-quark determinant is
a center-breaking source unless the coefficient \(h\) is set to zero.

\section{Controlled High-Temperature Holonomy Potential}
\label{sec:qcd-weiss-polyakov-holonomy-potential}

The Polyakov loop becomes a calculable order-coordinate in a regime where a
constant temporal holonomy is a valid background-field variable.  Choose a
spatially constant commuting background on
\(S^1_\beta\times\mathbb R^3\) and write its fundamental holonomy as
\begin{equation}
  P=\operatorname{diag}\left(e^{\ii\theta_1},\ldots,e^{\ii\theta_{N_c}}\right),
  \qquad
  \sum_{j=1}^{N_c}\theta_j\equiv0\pmod {2\pi}.
  \label{eq:qcd-weiss-holonomy-eigenphases}
\end{equation}
The variables \(\theta_j\) are defined modulo permutations and common shifts
compatible with \(\det P=1\).  A center transformation sends
\(P\mapsto zP\), \(z^{N_c}=1\).

\begin{controlledapproximation}[One-loop holonomy regime]
\label{ca:qcd-one-loop-holonomy-regime}
The one-loop holonomy potential is the leading weak-coupling thermal
effective potential for the zero Matsubara mode of \(A_0\), computed in a
background whose spatial gradients and noncommuting components are set to
zero.  Its expansion parameter is the running gauge coupling at scale
\(T\), and its infrared completion still requires the electric and magnetic
static sectors discussed in
Section~\ref{sec:qcd-phase-hard-thermal-loops-linde} below.
\end{controlledapproximation}

\paragraph{One-loop Polyakov-holonomy potential.}
For pure \(SU(N_c)\) Yang--Mills, the one-loop gluonic free-energy density
as a function of the holonomy \(P\) is
\begin{equation}
  f_g(P)
  =
  -\frac{2}{\pi^2\beta^4}
  \sum_{n=1}^{\infty}
  \frac{1}{n^4}\operatorname{tr}_{\rm adj}(P^n)
  =
  -\frac{2}{\pi^2\beta^4}
  \sum_{n=1}^{\infty}
  \frac{|\operatorname{tr}_{\square}P^n|^2-1}{n^4}.
  \label{eq:qcd-weiss-gluon-potential}
\end{equation}
For \(N_f\) massless Dirac fundamental quarks at \(\mu_q=0\), the one-loop
fermion contribution is
\begin{equation}
  f_q(P)
  =
  \frac{2N_f}{\pi^2\beta^4}
  \sum_{n=1}^{\infty}\frac{(-1)^n}{n^4}
  \left(
    \operatorname{tr}_{\square}P^n+
    \operatorname{tr}_{\square}P^{-n}
  \right).
  \label{eq:qcd-weiss-quark-potential}
\end{equation}

Consider first one bosonic oscillator with thermal phase \(\varphi\), meaning
that its Matsubara frequencies are shifted by the holonomy eigenvalue
\(e^{\ii\varphi}\).  Its thermal free-energy density is
\[
  T\int\frac{d^3p}{(2\pi)^3}
  \log\left(1-e^{-\beta |\mathbf p|+\ii\varphi}\right).
\]
Expanding the logarithm and integrating term by term gives
\begin{equation}
  -T\sum_{n\ge1}\frac{e^{\ii n\varphi}}{n}
  \int\frac{d^3p}{(2\pi)^3}e^{-n\beta|\mathbf p|}
  =
  -\frac{1}{\pi^2\beta^4}
  \sum_{n\ge1}\frac{e^{\ii n\varphi}}{n^4},
  \label{eq:qcd-weiss-boson-oscillator}
\end{equation}
because
\begin{equation}
  \int\frac{d^3p}{(2\pi)^3}e^{-n\beta|\mathbf p|}
  =
  \frac{1}{2\pi^2}\int_0^\infty p^2e^{-n\beta p}\,dp
  =
  \frac{1}{\pi^2n^3\beta^3}.
  \label{eq:qcd-weiss-thermal-momentum-integral}
\end{equation}
In background gauge, ghosts and unphysical vector polarizations cancel so
that two physical bosonic polarizations remain for each adjoint weight.
Summing \eqref{eq:qcd-weiss-boson-oscillator} over adjoint weights gives the
first expression in \eqref{eq:qcd-weiss-gluon-potential}.  The identity
\[
  \operatorname{tr}_{\rm adj}(P^n)
  =
  \operatorname{tr}_{\square}(P^n)
  \operatorname{tr}_{\overline\square}(P^n)-1
  =
  |\operatorname{tr}_{\square}P^n|^2-1
\]
uses \(\square\otimes\overline\square=\mathbf1\oplus{\rm adj}\).

For a Dirac fundamental quark at zero chemical potential the two spin
polarizations and the quark-antiquark pair give
\[
  -2T\int\frac{d^3p}{(2\pi)^3}
  \left[
    \log\left(1+e^{-\beta|\mathbf p|+\ii\theta_j}\right)
    +
    \log\left(1+e^{-\beta|\mathbf p|-\ii\theta_j}\right)
  \right]
\]
for each color eigenphase \(\theta_j\).  The plus sign inside the logarithm
is the antiperiodic thermal boundary condition.  Expanding
\(\log(1+x)=\sum_{n\ge1}(-1)^{n+1}x^n/n\), using
\eqref{eq:qcd-weiss-thermal-momentum-integral}, and summing over colors and
flavors gives \eqref{eq:qcd-weiss-quark-potential}.

\paragraph{Perturbative center-broken minima in pure Yang--Mills.}
For pure \(SU(N_c)\) Yang--Mills, the one-loop potential
\eqref{eq:qcd-weiss-gluon-potential} is minimized exactly at center
holonomies
\begin{equation}
  P=z\,\mathbf1_{N_c},
  \qquad
  z^{N_c}=1 .
  \label{eq:qcd-weiss-center-minima}
\end{equation}
At these minima
\begin{equation}
f_g(z\mathbf1)
  =
  -\frac{\pi^2T^4}{45}(N_c^2-1),
  \label{eq:qcd-weiss-broken-free-energy}
\end{equation}
which is the free gluon free-energy density.
For every unitary \(N_c\times N_c\) matrix,
\[
  |\operatorname{tr}_{\square}P^n|\le N_c .
\]
Hence \(\operatorname{tr}_{\rm adj}(P^n)\le N_c^2-1\) for every \(n\).  All
coefficients in \eqref{eq:qcd-weiss-gluon-potential} are negative, so the
minimum is obtained by maximizing each
\(\operatorname{tr}_{\rm adj}(P^n)\).  The \(n=1\) term already forces
\(|\operatorname{tr}_{\square}P|=N_c\), which for a unitary matrix means all
eigenvalues are equal.  The determinant-one condition then gives
\(P=z\mathbf1_{N_c}\) with \(z^{N_c}=1\).  Substitution in
\eqref{eq:qcd-weiss-gluon-potential} gives
\[
  f_g(z\mathbf1)
  =
  -\frac{2}{\pi^2\beta^4}(N_c^2-1)
  \sum_{n\ge1}\frac1{n^4}
  =
  -\frac{2}{\pi^2\beta^4}(N_c^2-1)\frac{\pi^4}{90},
\]
which is \eqref{eq:qcd-weiss-broken-free-energy}.

\begin{example}[Uniform center-symmetric holonomy]
\label{ex:qcd-weiss-center-symmetric-holonomy}
Let \(P_{\rm cs}\) have eigenvalues uniformly spaced around the unit circle,
with an overall phase chosen so that \(P_{\rm cs}\in SU(N_c)\).  Then
\[
  \operatorname{tr}_{\square}P_{\rm cs}^n=0
  \quad(N_c\nmid n),
  \qquad
  |\operatorname{tr}_{\square}P_{\rm cs}^n|^2=N_c^2
  \quad(N_c\mid n).
\]
Therefore
\begin{equation}
  f_g(P_{\rm cs})
  =
  \frac{\pi^2T^4}{45}\left(1-\frac1{N_c^2}\right),
  \qquad
  f_g(P_{\rm cs})-f_g(\mathbf1)
  =
  \frac{\pi^2T^4}{45}\left(N_c^2-\frac1{N_c^2}\right).
  \label{eq:qcd-weiss-center-symmetric-cost}
\end{equation}
The perturbative high-temperature gluon determinant therefore energetically
selects a center sector rather than the uniformly distributed holonomy.
\end{example}

\begin{remark}[Fundamental quarks as a center source]
\label{rem:qcd-weiss-fundamental-quark-center-source}
The gluon potential depends on \(P\) through adjoint traces and is invariant
under \(P\mapsto zP\).  The quark potential
\eqref{eq:qcd-weiss-quark-potential} contains fundamental traces:
\[
  \operatorname{tr}_{\square}(zP)^n=z^n\operatorname{tr}_{\square}P^n .
\]
Thus dynamical fundamental quarks supply an explicit center-breaking source.
At \(P=\mathbf1\), \eqref{eq:qcd-weiss-quark-potential} gives
\begin{equation}
  f_q(\mathbf1)
  =
  -\frac{7\pi^2T^4}{180}N_cN_f,
  \label{eq:qcd-weiss-free-quark-free-energy}
\end{equation}
the negative of the free massless Dirac-quark pressure in
the free-plasma calculation below.
\end{remark}

\section{Chiral Symmetry and the Dirac Spectrum}

Now set \(\mu_B=0\) and take \(N_f\) Dirac quarks with a common mass \(m\).
In the chiral limit \(m=0\), and ignoring the anomalous axial \(U(1)\), the
classical flavor symmetry is
\[
  SU(N_f)_L\times SU(N_f)_R\times U(1)_B .
\]
Let \(V_3\) be the spatial volume and let
\[
  \mathcal V_4=\beta V_3
\]
be the Euclidean thermal volume.  For degenerate quark mass, the pressure
derivative is the flavor-summed scalar source response.  We use the positive
per-flavor condensate convention
\begin{equation}
  \Sigma_m(T)
  =
  \frac1{N_f}\frac{\partial p(T,m)}{\partial m}.
  \label{eq:qcd-phase-chiral-condensate-source}
\end{equation}
Equivalently, after integrating the fermions in a fixed gauge background,
\(\Sigma_m(T)\) is the thermal average of
\(\mathcal V_4^{-1}\operatorname{Tr}(D_E+m)^{-1}\) for the one-flavor
Euclidean Dirac operator \(D_E\).  With the Grassmann-ordering convention in
which the local scalar density is
\(\bar q q=\sum_{i=1}^{N_f}\bar q_iq_i\), its flavor-summed expectation is
\[
  \langle\bar q q\rangle_{\mathrm{sum},T,m}
  =
  -N_f\Sigma_m(T).
\]
Thus \(\Sigma_m\ge0\) in a phase with the usual sign of the scalar condensate.
The chiral order parameter is the limit \(\Sigma(T)=\lim_{m\downarrow0}
\lim_{V_3\to\infty}\Sigma_m(T)\), when this limit exists.  A nonzero
condensate in the chiral limit transforms as
\((\overline{\mathbf N_f},\mathbf N_f)\oplus
(\mathbf N_f,\overline{\mathbf N_f})\) and breaks
\[
  SU(N_f)_L\times SU(N_f)_R
  \longrightarrow
  SU(N_f)_V .
\]

\begin{definition}[Chiral order parameter]
\label{def:qcd-chiral-order-parameter}
The chiral condensate used as an order parameter is
\begin{equation}
  \Sigma(T)
  =
  \lim_{m\downarrow0}\lim_{V_3\to\infty}
  \frac{1}{\mathcal V_4}
  \left\langle \operatorname{Tr}(D_E+m)^{-1}\right\rangle_{V_3,T,m},
  \label{eq:qcd-phase-sigma-definition}
\end{equation}
where \(D_E\) is the one-flavor massless Euclidean Dirac operator in a gauge
background at \(\mu_B=0\).  The infinite-volume limit precedes the chiral
limit.
\end{definition}

\paragraph{Banks--Casher relation under a spectral-density hypothesis.}
The nontrivial input is the existence of a thermodynamic Dirac spectral
density continuous at the origin.  Assume that the positive Dirac eigenvalue
density
\[
  \rho_{\mathcal V_4}(\lambda)
  =
  \frac1{\mathcal V_4}\left\langle
  \sum_{\lambda_n>0}\delta(\lambda-\lambda_n)
  \right\rangle
\]
has a thermodynamic limit \(\rho(\lambda)\) that is continuous at
\(\lambda=0\), and that zero-mode contributions have zero density in the
same limit.  Then
\begin{equation}
  \Sigma(T)=\pi\rho(0)
  \label{eq:qcd-phase-banks-casher}
\end{equation}
with the normalization in
\eqref{eq:qcd-phase-sigma-definition}.  If \(\rho(0)=0\), this particular
condensate vanishes.

At \(\mu_B=0\), \(D_E\) is anti-Hermitian in the usual positive Euclidean
inner product and anticommutes with \(\gamma_5\) in the massless theory.
Nonzero eigenvalues occur in pairs \(\pm\ii\lambda_n\), \(\lambda_n>0\).
For fixed positive mass,
\[
  \frac1{\mathcal V_4}\operatorname{Tr}(D_E+m)^{-1}
  =
  \frac{N_0}{\mathcal V_4m}
  +
  \frac1{\mathcal V_4}\sum_{\lambda_n>0}
  \left(\frac1{m+\ii\lambda_n}+\frac1{m-\ii\lambda_n}\right),
\]
where \(N_0\) is the number of zero modes.  The nonzero pair contribution is
\[
  \frac{2m}{\mathcal V_4}\sum_{\lambda_n>0}\frac1{\lambda_n^2+m^2}.
\]
By the zero-density hypothesis the first term drops out after
\(V_3\to\infty\).  The condensate definition therefore gives
\[
  \Sigma(T)
  =
  \lim_{m\downarrow0}
  \int_0^\infty
  \frac{2m\,\rho(\lambda)}{\lambda^2+m^2}\,d\lambda .
\]
The kernels
\[
  K_m(\lambda)=\frac{2m}{\lambda^2+m^2},\qquad \lambda\ge0,
\]
form an approximate identity on the half-line with total mass \(\pi\):
\[
  \int_0^\infty K_m(\lambda)\,d\lambda=\pi .
\]
Continuity of \(\rho\) at the origin gives the limit
\(\pi\rho(0)\), and hence \eqref{eq:qcd-phase-banks-casher}.  If instead
one defines the spectral density on all of \(\mathbb R\), counting both
positive and negative eigenvalues, the same formula holds with that
two-sided density evaluated at zero.

\begin{remark}[Order of limits]
The relation is a statement about the thermodynamic spectral density near
zero.  Setting \(m=0\) at finite volume first gives no condensate from paired
nonzero eigenvalues.  The density of eigenvalues at spacing
\(O(\mathcal V_4^{-1})\) near the origin is precisely the infinite-volume
order parameter that can survive the subsequent chiral limit.
\end{remark}

\subsection{Axial Ward Identity and Thermal GMOR Logic}
\label{subsec:qcd-thermal-axial-ward-gmor}

The condensate is tied to integrated pseudoscalar correlations by an axial
Ward identity.  This is the
cleanest finite-temperature form of the Gell-Mann--Oakes--Renner logic,
because at \(T>0\) Lorentz invariance is absent and one must specify which
pion pole, screening pole, or spectral residue is being used.

Let \(\tau^a\), \(a=1,\ldots,N_f^2-1\), be Hermitian generators of
\(\mathfrak{su}(N_f)\) normalized by
\begin{equation}
  \operatorname{tr}_{f}(\tau^a\tau^b)=\delta^{ab}.
  \label{eq:qcd-flavor-generator-normalization}
\end{equation}
Define the nonsinglet Euclidean axial current and pseudoscalar density by
\begin{equation}
  A_\mu^a=\bar q\gamma_\mu\gamma_5\tau^a q,
  \qquad
  P^a=\bar q\,\ii\gamma_5\tau^a q .
  \label{eq:qcd-axial-current-pseudoscalar-density}
\end{equation}
The factor of \(\ii\) in \(P^a\) is part of the Euclidean source convention:
a real pseudoscalar source \(p^a\) appears in the mass matrix as
\(m\mathbf1+\ii\gamma_5p^a\tau^a\).  With this convention the Euclidean
pseudoscalar susceptibility is
\begin{equation}
  \chi_\pi^{ab}(T,m)
  =
  -\int_0^\beta d\tau\int_{\mathbb R^3}d^3x\,
  \langle P^a(\tau,\mathbf x)P^b(0)\rangle_{c,T,m},
  \label{eq:qcd-pseudoscalar-susceptibility-definition}
\end{equation}
with the finite-volume definition obtained by replacing \(\mathbb R^3\) by
the spatial torus and taking the thermodynamic limit after the integral.

\begin{lemma}[Integrated nonsinglet axial Ward identity]
\label{lem:qcd-integrated-axial-ward}
Assume a regulator and renormalization prescription preserving the nonsinglet
vector flavor symmetry and the nonsinglet axial Ward identity up to the
explicit mass term; there is no nonsinglet axial anomaly.  Then
\begin{equation}
  2m\,\chi_\pi^{ab}(T,m)
  =
  -\left\langle
  \bar q\{\tau^a,\tau^b\}q
  \right\rangle_{T,m}.
  \label{eq:qcd-integrated-axial-ward}
\end{equation}
In a flavor-symmetric thermal state this becomes
\begin{equation}
  m\,\chi_\pi^{ab}(T,m)=\delta^{ab}\Sigma_m(T).
  \label{eq:qcd-flavor-symmetric-pseudoscalar-ward}
\end{equation}
\end{lemma}

\begin{proof}
The proof is a finite-regulator Ward-identity argument, followed by the
thermodynamic limit in the order specified in the definition of the
susceptibility.  Perform the local nonsinglet axial change of variables
\[
  \delta q=\ii\alpha^a(x)\tau^a\gamma_5 q,\qquad
  \delta\bar q=\ii\alpha^a(x)\bar q\gamma_5\tau^a .
\]
The regulator hypothesis says that the fermion measure has no nonsinglet
Jacobian anomaly.  The variation of the mass term is fixed by the Euclidean
source convention in \eqref{eq:qcd-axial-current-pseudoscalar-density}:
\[
  \delta_\alpha\bigl(m\bar q q\bigr)
  =
  2m\,\alpha^a\,\bar q\ii\gamma_5\tau^a q
  =
  2m\,\alpha^a P^a .
\]
The kinetic term gives the divergence of the current \(A_\mu^a\).  With the
active-variation convention used here, differentiating the change-of-variables
identity with respect to \(\alpha^a(x)\) and applying it to the insertion
\(P^b(0)\) gives the distributional identity
\begin{equation}
  \partial_\mu
  \langle A_\mu^a(x)P^b(0)\rangle_{T,m}
  +2m\langle P^a(x)P^b(0)\rangle_{T,m}
  +\delta^{(4)}_\beta(x)
  \langle \delta^a P^b(0)\rangle_{T,m}
  =
  0 .
  \label{eq:qcd-local-axial-ward-pseudoscalar}
\end{equation}
Here \(\delta^{(4)}_\beta\) is the periodic delta distribution on the thermal
spacetime torus before the thermodynamic limit, and
\(\delta^aP^b\) denotes the variation of the inserted operator, not a second
source variation.  The contact term is
\[
  \delta^a P^b(0)
  =
  \ii\,\bar q\gamma_5\tau^a\,\ii\gamma_5\tau^b q
  +
  \bar q\,\ii\gamma_5\tau^b\,\ii\tau^a\gamma_5 q
  =
  -\bar q\{\tau^a,\tau^b\}q(0).
\]
Integrating \eqref{eq:qcd-local-axial-ward-pseudoscalar} over the thermal
spacetime torus removes the divergence term by periodicity of the bosonic
current.  Vector flavor invariance gives \(\langle P^a\rangle_{T,m}=0\) for
nonsinglet \(a\), so the connected and ordinary two-point functions coincide
in this channel.  Using the susceptibility definition
\eqref{eq:qcd-pseudoscalar-susceptibility-definition}, one obtains
\[
  2m\,\chi_\pi^{ab}(T,m)
  =
  \langle\delta^a P^b\rangle_{T,m}
  =
  -
  \left\langle
    \bar q\{\tau^a,\tau^b\}q
  \right\rangle_{T,m},
\]
which is \eqref{eq:qcd-integrated-axial-ward}.  No statement about a pion pole
has entered; the pole-saturation input is a separate infrared hypothesis.

In a flavor-symmetric state
\[
  \left\langle\bar q\{\tau^a,\tau^b\}q\right\rangle_{T,m}
  =
  \frac{2}{N_f}\delta^{ab}
  \langle\bar q q\rangle_{\mathrm{sum},T,m}
  =
  -2\delta^{ab}\Sigma_m(T),
\]
where \eqref{eq:qcd-flavor-generator-normalization} fixes the singlet part of
the anticommutator.  Substitution gives
\eqref{eq:qcd-flavor-symmetric-pseudoscalar-ward}.
\end{proof}

\paragraph{GMOR as a pole-saturation consequence.}
The integrated Ward identity becomes the finite-temperature GMOR relation
only after one supplies a spectral hypothesis about the small-mass broken
phase.  Suppose, in addition to the hypotheses of
Lemma~\ref{lem:qcd-integrated-axial-ward}, that for small \(m>0\) the
zero-spatial-momentum pseudoscalar susceptibility in a broken phase is
saturated at leading order by an isolated pion pole of mass \(M_\pi(T,m)\).
Let \(G_\pi(T,m)\) and \(f_{t,\delta}(T,m)\) be defined in the same flavor
normalization \eqref{eq:qcd-flavor-generator-normalization} by the pole
matrix elements
\[
  \langle\Omega_T|P^a(0)|\pi^b\rangle
  =
  \delta^{ab}G_\pi,
  \qquad
  \langle\Omega_T|A_0^a(0)|\pi^b\rangle
  =
  \delta^{ab}f_{t,\delta}M_\pi,
\]
or by the corresponding KMS spectral residues when a thermal Hilbert-space
representative is used.  If the pole obeys the axial Ward relation
\begin{equation}
  2m\,G_\pi=f_{t,\delta}M_\pi^2
  \label{eq:qcd-pcac-pole-relation}
\end{equation}
and the non-pole part of \(\chi_\pi^{aa}\) is \(O(1)\) as
\(m\downarrow0\), then pole saturation gives
\[
  \chi_\pi^{ab}(T,m)
  =
  \delta^{ab}\frac{G_\pi(T,m)^2}{M_\pi(T,m)^2}+O(1).
\]
The Ward identity \eqref{eq:qcd-flavor-symmetric-pseudoscalar-ward} gives the
same susceptibility as
\[
  \chi_\pi^{ab}(T,m)
  =
  \delta^{ab}\frac{\Sigma(T)}{m}+o(m^{-1})
\]
when \(\Sigma_m(T)\to\Sigma(T)\).  Combining these two leading terms with
\eqref{eq:qcd-pcac-pole-relation},
\[
  \frac{G_\pi^2}{M_\pi^2}
  =
  \frac{f_{t,\delta}^2M_\pi^2}{4m^2},
\]
and equating the coefficients of \(1/m\) gives
\begin{equation}
  f_{t,\delta}(T,m)^2M_\pi(T,m)^2
  =
  4m\Sigma(T)+o(m).
  \label{eq:qcd-thermal-gmor-delta-normalization}
\end{equation}
If instead one uses generators \(t^a\) normalized by
\(\operatorname{tr}_{f}(t^at^b)=\frac12\delta^{ab}\), the decay constant is
smaller by \(\sqrt2\) and the same statement is
\begin{equation}
  f_{t,1/2}(T,m)^2M_\pi(T,m)^2
  =
  2m\Sigma(T)+o(m).
  \label{eq:qcd-thermal-gmor-half-normalization}
\end{equation}
Here rescaling the generators from
\(\operatorname{tr}_f(\tau^a\tau^b)=\delta^{ab}\) to
\(\operatorname{tr}_f(t^at^b)=\frac12\delta^{ab}\) rescales both the current
and the pseudoscalar density by \(1/\sqrt2\), hence
\(f_{t,1/2}=f_{t,\delta}/\sqrt2\), giving
\eqref{eq:qcd-thermal-gmor-half-normalization}.

\subsection{Low-Temperature Chiral Effective Theory}
\label{subsec:qcd-low-temperature-chiral-effective-theory}

The Ward identity identifies the source derivative of the pressure with the
integrated pion channel.  In the low-temperature phase this identity can be
evaluated in a controlled effective theory.  The control assumptions are
part of the statement: the spatial thermodynamic limit has been taken, the
temperature is small compared with the first non-Goldstone mass scale, and
the derivative expansion is organized at momenta \(p\ll 4\pi F\).

\begin{definition}[Leading chiral effective theory]
\label{def:qcd-leading-chiral-effective-theory}
For \(N_f\) degenerate light quarks, the leading Euclidean chiral effective
theory consists of a field \(U:\mathbb R^3\times S^1_\beta\to SU(N_f)\), a
constant \(F>0\), and a constant \(B>0\), with leading action
\begin{equation}
  S_{\chi}^{(2)}
  =
  \int_0^\beta d\tau\int d^3x
  \left[
    \frac{F^2}{4}
    \operatorname{tr}_f(\partial_\mu U^\dagger\partial_\mu U)
    -
    \frac{F^2B}{2}
    \operatorname{tr}_f\bigl(MU^\dagger+UM^\dagger\bigr)
  \right].
  \label{eq:qcd-leading-chiral-effective-action}
\end{equation}
Here \(M=m\mathbf1_{N_f}\) for the degenerate-mass discussion, and
\(\operatorname{tr}_f\) is the ordinary fundamental flavor trace.  We
parameterize the pion fields by
\begin{equation}
  U=\exp\left(\frac{\ii\sqrt2}{F}\pi^a\tau^a\right),
  \qquad
  \operatorname{tr}_f(\tau^a\tau^b)=\delta^{ab}.
  \label{eq:qcd-chiral-field-parameterization}
\end{equation}
With this convention \(F\) is the decay constant in the half-trace flavor
normalization; the trace-delta decay constant used in the pole-saturation
discussion above is
\(f_{t,\delta}=\sqrt2F+O(m,T^2)\).
\end{definition}

\begin{remark}[Tree-level source normalization]
\label{prop:qcd-chiral-effective-source-normalization}
The action \eqref{eq:qcd-leading-chiral-effective-action} gives
\begin{equation}
  \Sigma=F^2B,
  \qquad
  M_\pi^2=2Bm
  \label{eq:qcd-chiral-sigma-mpi-tree}
\end{equation}
at tree level, where \(\Sigma\) is the positive per-flavor condensate in
\eqref{eq:qcd-phase-chiral-condensate-source}.  Equivalently,
\[
  F^2M_\pi^2=2m\Sigma,
  \qquad
  f_{t,\delta}^2M_\pi^2=4m\Sigma
\]
at this order.
For a constant field \(U=\mathbf1\), the Euclidean free-energy density from
\eqref{eq:qcd-leading-chiral-effective-action} is
\[
  f_0(m)=-N_fF^2Bm,
  \qquad
  p_0(m)=N_fF^2Bm .
\]
Therefore \(N_f^{-1}\partial p_0/\partial m=F^2B\), proving
\(\Sigma=F^2B\).  Expanding
\eqref{eq:qcd-chiral-field-parameterization},
\[
  U+U^\dagger
  =
  2\mathbf1-\frac{2}{F^2}\pi^a\pi^b\tau^a\tau^b+O(\pi^4),
\]
and substituting \(M=m\mathbf1\) gives the quadratic Euclidean Lagrangian
\[
  \frac12\partial_\mu\pi^a\partial_\mu\pi^a
  +
  Bm\,\pi^a\pi^a .
\]
Comparing with
\(\frac12(\partial\pi)^2+\frac12M_\pi^2\pi^2\) gives
\(M_\pi^2=2Bm\).  The two GMOR forms follow from
\(f_{t,\delta}=\sqrt2F\) at leading order.
\end{remark}

\begin{lemma}[One-loop pion-gas correction to the condensate]
\label{lem:qcd-low-temperature-chiral-condensate}
Under the leading chiral effective assumptions above, and in the chiral limit
after the spatial thermodynamic limit, the low-temperature condensate obeys
\begin{equation}
  \frac{\Sigma(T)}{\Sigma(0)}
  =
  1-\frac{N_f^2-1}{12N_f}\frac{T^2}{F^2}
  +O\!\left(\frac{T^4}{F^4}\right)
  \label{eq:qcd-low-temperature-chiral-condensate}
\end{equation}
within the chiral derivative expansion.  For \(N_f=2\), this is
\[
  \Sigma(T)/\Sigma(0)=1-\frac{T^2}{8F^2}+O(T^4/F^4).
\]
\end{lemma}

\begin{proof}
To one loop in the leading effective theory the only \(m\)-dependent thermal
term at order \(T^2\) is the free pressure of
\(d_\pi=N_f^2-1\) pion fields with
\(\omega_k=(k^2+M_\pi^2)^{1/2}\):
\begin{equation}
  p_\pi(T,M_\pi)
  =
  -d_\pi T\int\frac{d^3k}{(2\pi)^3}
  \log\bigl(1-e^{-\beta\omega_k}\bigr).
  \label{eq:qcd-free-pion-pressure}
\end{equation}
The vacuum part is absorbed into the zero-temperature low-energy constants;
the displayed thermal part is finite.  Differentiating at fixed \(T\),
\begin{equation}
  \frac{\partial p_\pi}{\partial M_\pi^2}
  =
  -\frac{d_\pi}{2}
  \int\frac{d^3k}{(2\pi)^3}
  \frac{1}{\omega_k}\frac{1}{e^{\beta\omega_k}-1}.
  \label{eq:qcd-pion-pressure-mass-derivative}
\end{equation}
Since \(M_\pi^2=2Bm\), the pion contribution to the per-flavor condensate is
\[
  \frac1{N_f}\frac{\partial p_\pi}{\partial m}
  =
  -\frac{d_\pi B}{N_f}
  \int\frac{d^3k}{(2\pi)^3}
  \frac{1}{\omega_k}\frac{1}{e^{\beta\omega_k}-1}.
\]
In the chiral limit the integral is
\begin{equation}
  \int\frac{d^3k}{(2\pi)^3}
  \frac{1}{|k|}\frac{1}{e^{\beta |k|}-1}
  =
  \frac{T^2}{12}.
  \label{eq:qcd-massless-pion-tadpole}
\end{equation}
Thus
\[
  \Sigma(T)
  =
  F^2B-\frac{N_f^2-1}{N_f}B\,\frac{T^2}{12}
  +O(T^4/F^2).
\]
Dividing by \(\Sigma(0)=F^2B\) gives
\eqref{eq:qcd-low-temperature-chiral-condensate}.  The next terms require
the \(O(p^4)\) chiral Lagrangian and two-loop pion interactions; they are
not fixed by symmetry and \(F,B\) alone.
\end{proof}

\begin{remark}[What the low-temperature formula establishes]
Equation~\eqref{eq:qcd-low-temperature-chiral-condensate} is a controlled
asymptotic statement inside the broken phase.  It does not by itself prove a
chiral restoration temperature or determine the order of a transition.  Such
questions require the nonperturbative thermal QCD limit, or an effective
description whose domain includes the critical region.
\end{remark}

\section{Free Plasma Limit and Thermodynamic Observables}

At asymptotically high temperature and fixed \(\mu_B/T\), asymptotic freedom
makes the leading term in the pressure the free massless quark-gluon pressure.
It is the leading term in a controlled high-temperature expansion whose
subleading terms require screening and infrared resummation.

\paragraph{Massless free QCD pressure.}
For \(SU(N_c)\) with \(N_f\) massless Dirac fundamental quarks at
\(\mu_B=0\), the free pressure is
\begin{equation}
  p_{\rm SB}(T)
  =
  \frac{\pi^2T^4}{45}(N_c^2-1)
  +
  \frac{7\pi^2T^4}{180}N_cN_f .
  \label{eq:qcd-phase-stefan-boltzmann}
\end{equation}
At nonzero baryon chemical potential, the quark part becomes
\begin{equation}
  p_{q,{\rm free}}(T,\mu_B)
  =
  N_cN_f\left[
  \frac{7\pi^2T^4}{180}
  +\frac{\mu_B^2T^2}{6N_c^2}
  +\frac{\mu_B^4}{12\pi^2N_c^4}
  \right].
  \label{eq:qcd-phase-free-quark-mu-pressure}
\end{equation}

A massless bosonic degree of freedom contributes
\[
  -T\int\frac{d^3p}{(2\pi)^3}\log(1-e^{-|p|/T})
  =
  \frac{\pi^2T^4}{90},
\]
using \(\int_0^\infty x^3(e^x-1)^{-1}dx=\pi^4/15\).  A gluon has two
physical polarizations and \(N_c^2-1\) color states, giving the first term.

A massless fermionic degree of freedom contributes \(7/8\) times the bosonic
answer at zero chemical potential.  A Dirac quark has four
particle-antiparticle and spin degrees for each color and flavor, giving
\((7/8)(4N_cN_f)\pi^2T^4/90\), which is the second term of
\eqref{eq:qcd-phase-stefan-boltzmann}.  The standard Fermi integral with
chemical potential \(\mu_q\) gives, for one Dirac quark of one color,
\[
  \frac{7\pi^2T^4}{180}
  +\frac{\mu_q^2T^2}{6}
  +\frac{\mu_q^4}{12\pi^2}.
\]
Substituting \(\mu_q=\mu_B/N_c\) and multiplying by \(N_cN_f\) gives
\eqref{eq:qcd-phase-free-quark-mu-pressure}.

The energy density, entropy density, and trace anomaly are then defined by
\[
  s=\frac{\partial p}{\partial T},\qquad
  n_B=\frac{\partial p}{\partial\mu_B},\qquad
  \epsilon=-p+Ts+\mu_B n_B,
\]
and
\[
  \Delta(T,\mu_B)=\epsilon-3p .
\]
For a conformal thermal state in four dimensions, \(\Delta=0\).  In QCD,
\(\Delta\) measures the breaking of scale invariance by quark masses, by the
running coupling, and by the thermal state itself.

\begin{remark}[Thermodynamic identities]
\label{rem:qcd-thermodynamic-derivative-identities}
Let \(p(T,\mu_B)\) be the thermodynamic pressure of a QCD phase on an open
domain in which it is twice continuously differentiable.  Define
\[
  s=\partial_Tp,\qquad n_B=\partial_{\mu_B}p,\qquad
  \epsilon=-p+Ts+\mu_Bn_B .
\]
Then
\[
  dp=s\,dT+n_B\,d\mu_B,
\]
and
\[
  d\epsilon=T\,ds+\mu_B\,dn_B,
\]
and the interaction measure is
\[
  \Delta(T,\mu_B)=\epsilon-3p
  =
  T\partial_Tp+\mu_B\partial_{\mu_B}p-4p .
\]
Equivalently, at fixed \(x=\mu_B/T\),
\[
  \frac{\Delta(T,\mu_B)}{T^4}
  =
  T\,\partial_T\left(\frac{p(T,xT)}{T^4}\right)_{x}.
\]
At \(\mu_B=0\), if charge conjugation gives \(n_B(T,0)=0\) and
\(\partial_T\epsilon(T,0)>0\), the adiabatic sound speed determined by the
pressure is
\[
  c_s^2(T)
  =
  \frac{dp(T,0)/dT}{d\epsilon(T,0)/dT}
  =
  \frac{\partial_Tp(T,0)}{T\,\partial_T^2p(T,0)} .
\]
These formulae are definitions and thermodynamic consequences of the
pressure; they are not model assumptions about quasiparticles.

The first identity is the definition of \(s\) and \(n_B\) as derivatives of
the limiting pressure.  Differentiating
\(\epsilon=-p+Ts+\mu_Bn_B\) gives
\[
  d\epsilon
  =
  -dp+T\,ds+s\,dT+\mu_B\,dn_B+n_B\,d\mu_B.
\]
Substituting \(dp=s\,dT+n_B\,d\mu_B\) leaves
\(d\epsilon=T\,ds+\mu_B\,dn_B\).  The displayed formula for \(\Delta\)
follows by substituting \(s=\partial_Tp\) and
\(n_B=\partial_{\mu_B}p\) into \(\epsilon-3p\).

For the fixed-\(x\) identity, write \(\mu_B=xT\).  The chain rule gives
\[
  T\,\partial_T\left(\frac{p(T,xT)}{T^4}\right)_{x}
  =
  \frac{T\partial_Tp+\mu_B\partial_{\mu_B}p-4p}{T^4}.
\]
At \(\mu_B=0\), charge conjugation removes the baryon-density term.  Along
that one-parameter equilibrium curve,
\[
  \frac{d\epsilon}{dT}
  =
  T\,\frac{ds}{dT}
  =
  T\,\partial_T^2p(T,0),
  \qquad
  \frac{dp}{dT}=\partial_Tp(T,0),
\]
which gives the sound-speed formula whenever the denominator is nonzero.
\end{remark}

\begin{example}[Conformal Stefan--Boltzmann benchmark]
\label{ex:qcd-conformal-sb-thermodynamics}
If
\[
  p(T,\mu_B)=aT^4+b\mu_B^2T^2+c\mu_B^4
\]
with \(a,b,c\) independent of \(T\) and \(\mu_B\), then \(p\) is homogeneous
of degree four, so Euler's identity gives
\[
  T\partial_Tp+\mu_B\partial_{\mu_B}p=4p,\qquad \Delta=0.
\]
At \(\mu_B=0\), one obtains \(\epsilon=3aT^4\) and
\[
  c_s^2=\frac{4aT^3}{12aT^3}=\frac13 .
\]
The free plasma is therefore a normalization benchmark for lattice and
resummed calculations of \(\Delta/T^4\) and \(c_s^2\); departures from it
measure interactions, masses, running couplings, and nonperturbative thermal
physics.
\end{example}

\section{Hard Thermal Loops and the Linde Obstruction}
\label{sec:qcd-phase-hard-thermal-loops-linde}

Chapter~\ref{ch:thermal-gauge-theory-screening} derived the one-loop Debye
mass in the present trace convention:
\[
  m_D^2
  =
  g^2T^2\left(
  \frac{C_A}{3}
  +\frac{1}{3}\sum_{\text{Dirac }f}T_{R_f}
  \right)
\]
for QCD without scalars.  For \(SU(N_c)\) with
\(\operatorname{tr}_{\square}(T^aT^b)=\delta^{ab}\), this gives
\[
  C_A=2N_c,\qquad T_{\square}=1.
\]
The scale \(gT\) is electric.  It must be resummed in static longitudinal
propagators before a weak-coupling expansion of many thermal quantities is
well-defined.

\begin{controlledapproximation}[Hard-thermal-loop regime]
\label{ca:qcd-htl-regime}
The HTL approximation applies to external frequencies and momenta
\[
  \omega,\ |\mathbf k| = O(gT)
\]
in a plasma whose internal loop momenta are hard, \(p=O(T)\).  The leading
soft gauge-field effective action is the gauge-invariant functional whose
second variation reproduces the transverse and longitudinal one-loop
polarization tensors in this kinematic regime.  A calculation outside this
scale separation is not an HTL calculation.
\end{controlledapproximation}

\paragraph{Static HTL polarization and Debye mass.}
Let \(a_0=a_0^aT^a\) be a slowly varying canonically normalized temporal
gauge potential, and normalize the generators by
\[
  \operatorname{tr}_R(T_R^aT_R^b)=T_R\,\delta^{ab},\qquad
  f^{acd}f^{bcd}=C_A\,\delta^{ab}.
\]
For a weakly coupled thermal plasma at \(\mu_B=0\), the static HTL color
susceptibility is
\[
  \Pi_{00}^{ab}(0,\mathbf k\to0)
  =
  \delta^{ab}m_D^2,
  \qquad
  m_D^2
  =
  g^2T^2\left(\frac{C_A}{3}
  +\frac{1}{3}\sum_{\text{Dirac }f}T_{R_f}\right).
\]
In the present \(SU(N_c)\) convention
\(\operatorname{tr}_{\square}(T^aT^b)=\delta^{ab}\), one has
\[
  C_A=2N_c,\qquad T_{\square}=1,
\]
and hence
\[
  m_D^2=g^2T^2\left(\frac{2N_c}{3}+\frac{N_f}{3}\right).
\]
To leading static order the longitudinal electric propagator is screened as
\[
  D_{00}^{ab}(0,\mathbf k)
  =
  \frac{\delta^{ab}}{\mathbf k^2+m_D^2},
\]
with the understanding that this formula is the static electric HTL limit,
not a gauge-invariant definition of a particle pole.

The statement is a color-susceptibility calculation for hard particles.
Let a hard excitation of momentum \(p=|\mathbf p|\) and color weight \(w\)
couple to the soft potential by the energy shift
\(-g a_0^aw^a\).  Linearizing its distribution gives
\[
  \delta n(p)=g a_0^bw^b\,[-n'(p)] .
\]
Multiplying by the color charge \(gw^a\), summing over weights, and
integrating over hard momenta gives the induced color density.  Therefore a
species in representation \(R\) contributes the positive susceptibility
\[
  g^2\,d_{\mathrm{spin/particle}}\,
  \operatorname{tr}_R(T_R^aT_R^b)
  \int \frac{d^3p}{(2\pi)^3}\,[-n'(p)] .
\]
For gluons the representation is the adjoint and there are two transverse
hard polarizations.  For a massless Dirac fermion there are two spin states
and independent particle and antiparticle excitations; the antiparticle
charge is opposite, but the susceptibility is quadratic in the charge, so
the total degeneracy factor is \(4\).

The required thermal integrals are elementary but convention-sensitive:
\[
  I_B
  =
  \int \frac{d^3p}{(2\pi)^3}\,[-n_B'(p)]
  =
  \frac{1}{\pi^2}\int_0^\infty dp\,p\,n_B(p)
  =\frac{T^2}{6},
\]
\[
  I_F
  =
  \int \frac{d^3p}{(2\pi)^3}\,[-n_F'(p)]
  =
  \frac{1}{\pi^2}\int_0^\infty dp\,p\,n_F(p)
  =\frac{T^2}{12}.
\]
The integration by parts has no boundary term: at \(p=\infty\) the
distributions decay exponentially, while at \(p=0\) one has
\(p^2n_B(p)\to0\) and \(p^2n_F(p)\to0\); moreover
\(p^2[-n'_B(p)]\) and \(p^2[-n'_F(p)]\) are locally integrable.  The last
equalities use
\[
  \int_0^\infty \frac{x\,dx}{e^x-1}=\frac{\pi^2}{6},
  \qquad
  \int_0^\infty \frac{x\,dx}{e^x+1}=\frac{\pi^2}{12}.
\]
Thus
\[
  m_D^2
  =
  g^2\bigl(2C_A I_B+4\sum_fT_{R_f}I_F\bigr)
  =
  g^2T^2\left(\frac{C_A}{3}
  +\frac{1}{3}\sum_fT_{R_f}\right).
\]
The resummed static longitudinal propagator is the geometric series obtained
by inserting this static self-energy into the soft electric propagator.  The
derivation has used only the HTL separation between hard loop momentum and
soft external field; it does not determine the magnetic sector.

\begin{definition}[Retarded HTL induced-current kernel]
\label{def:qcd-retarded-htl-induced-current-kernel}
Work in the rest frame of the equilibrium state, with mostly-plus Lorentzian
metric and \(K=(\omega,\mathbf k)\).  Let
\[
  v^\mu=(1,\mathbf v),\qquad |\mathbf v|=1,\qquad
  \int_{\mathbf v}(\cdots)
  =
  \int_{S^2}\frac{\dd\Omega_{\mathbf v}}{4\pi}(\cdots).
\]
The retarded HTL induced-current kernel is the distribution
\(\mathcal K_R^{\mu\nu,ab}(\omega,\mathbf k)\) defined by
\[
  j^\mu_a(\omega,\mathbf k)
  =
  \mathcal K_R^{\mu\nu,ab}(\omega,\mathbf k)A_\nu^b(\omega,\mathbf k),
\]
where
\[
  \mathcal K_R^{\mu\nu,ab}(\omega,\mathbf k)
  =
  \delta^{ab}m_D^2
  \left[
    \delta^\mu_0\delta^\nu_0
    -
    \omega\int_{\mathbf v}
    \frac{v^\mu v^\nu}
    {\omega-\mathbf v\cdot\mathbf k+i0}
  \right].
\]
The \(+i0\) prescription is part of the definition: it is the retarded
boundary condition for the collisionless response of hard particles.  This
kernel is normalized so that
\[
  \mathcal K_R^{00,ab}(0,\mathbf k)=\delta^{ab}m_D^2,
\]
matching the static HTL calculation above.
\end{definition}

\paragraph{Gauge-covariant HTL equations and transversality.}
Let \(W_a(x,\mathbf v)\) be an adjoint auxiliary field on spacetime times the
unit velocity sphere.  The gauge-covariant linear-response equations
\[
  (v\cdot D)W(x,\mathbf v)=v^iF_{i0}(x),\qquad
  j^\mu(x)=m_D^2\int_{\mathbf v}v^\mu W(x,\mathbf v),
\]
where \(D\) is the adjoint covariant derivative, are covariantly conserved:
\[
  D_\mu j^\mu=0.
\]
Linearizing around \(A_\mu=0\) and imposing the retarded prescription gives
the kernel of Definition~\ref{def:qcd-retarded-htl-induced-current-kernel}.
Consequently, in momentum space,
\[
  K_\mu\mathcal K_R^{\mu\nu,ab}(K)=0,
  \qquad
  K_\mu=(-\omega,\mathbf k),
\]
so the HTL response is transverse in the mostly-plus convention.

The verification is explicit.
The auxiliary field \(W\) transforms in the adjoint representation.  Applying
the covariant divergence to the current gives
\[
  D_\mu j^\mu
  =
  m_D^2\int_{\mathbf v}(v\cdot D)W
  =
  m_D^2\int_{\mathbf v}v^iF_{i0}.
\]
Since \(F_{i0}\) is independent of \(\mathbf v\) and
\(\int_{\mathbf v}v^i=0\), the right side vanishes.  This proves covariant
current conservation before choosing a gauge.

For the linearized kernel, take fields proportional to
\(\exp(-i\omega t+i\mathbf k\cdot\mathbf x)\).  The retarded solution of
\((\partial_t+\mathbf v\cdot\nabla)W=v^iF_{i0}\) is
\[
  W(K,\mathbf v)
  =
  -\,\frac{\mathbf v\cdot\mathbf k\,A_0(K)
  +\omega\,v^iA_i(K)}
  {\omega-\mathbf v\cdot\mathbf k+i0},
\]
using \(F_{i0}=\partial_iA_0-\partial_0A_i\).  Inserting this solution into
\(j^\mu=m_D^2\int_{\mathbf v}v^\mu W\) and using
\(\int_{\mathbf v}v^i=0\) gives
\[
  j^\mu
  =
  m_D^2\left[
    \delta^\mu_0\delta^\nu_0
    -
    \omega\int_{\mathbf v}
    \frac{v^\mu v^\nu}
    {\omega-\mathbf v\cdot\mathbf k+i0}
  \right]A_\nu,
\]
which is precisely the stated kernel, with the color factor
\(\delta^{ab}\) restored.

Transversality is an algebraic consequence of the same formula.  With
\(K_\mu=(-\omega,\mathbf k)\),
\[
  K_\mu v^\mu=-(\omega-\mathbf v\cdot\mathbf k).
\]
Therefore
\[
  K_\mu\mathcal K_R^{\mu\nu}
  =
  m_D^2\left[
    -\omega\,\delta^\nu_0
    +
    \omega\int_{\mathbf v}v^\nu
  \right].
\]
The integral of \(v^0\) is \(1\), and the integral of \(v^i\) is \(0\).
Thus the expression vanishes for \(\nu=0\) and for each spatial \(\nu=i\).

\paragraph{Longitudinal HTL response and Landau cut.}
For \(k=|\mathbf k|>0\),
\[
  \mathcal K_R^{00,ab}(\omega,\mathbf k)
  =
  \delta^{ab}m_D^2
  \left[
    1-\frac{\omega}{2k}
    \log\frac{\omega+k+i0}{\omega-k+i0}
  \right].
\]
Equivalently,
\[
  \int_{\mathbf v}
  \frac{1}{\omega-\mathbf v\cdot\mathbf k+i0}
  =
  \frac{1}{2k}
  \log\frac{\omega+k+i0}{\omega-k+i0}.
\]
For \(|\omega|<k\), the logarithm has an imaginary part.  This is the
Landau cut of the retarded thermal response, caused by hard particles whose
velocity satisfies \(\omega=\mathbf v\cdot\mathbf k\).

Choose the \(z\)-axis along \(\mathbf k\) and write
\(\mu=\cos\theta\).  Then
\[
  \int_{\mathbf v}
  \frac{1}{\omega-\mathbf v\cdot\mathbf k+i0}
  =
  \frac12\int_{-1}^{1}
  \frac{\dd\mu}{\omega-k\mu+i0}.
\]
The antiderivative is \(-k^{-1}\log(\omega-k\mu+i0)\).  Evaluating it at
\(\mu=1\) and \(\mu=-1\) gives the displayed logarithm.  Substitution in
Definition~\ref{def:qcd-retarded-htl-induced-current-kernel} gives
\(\mathcal K_R^{00}\).

The distribution identity
\[
  \frac{1}{\omega-k\mu+i0}
  =
  \operatorname{PV}\frac{1}{\omega-k\mu}
  -i\pi\delta(\omega-k\mu)
\]
shows that an imaginary part is present exactly when the equation
\(\mu=\omega/k\) has a solution in \([-1,1]\), namely when
\(|\omega|<k\).  This proves the Landau-cut statement and identifies its
kinematic origin.

\begin{remark}[Magnetic sector at \(g^6T^4\)]
\label{rem:qcd-linde-scale}
In four-dimensional high-temperature Yang--Mills theory, the static magnetic
sector has three-dimensional coupling
\[
  g_3^2=g^2T .
\]
If the magnetic sector develops a mass gap of order \(g_3^2\), its
contribution to the four-dimensional pressure begins at order
\[
  T(g_3^2)^3 = g^6T^4 .
\]
The coefficient of this contribution is not determined by ordinary
perturbation theory around the massless magnetic theory.

After the nonzero Matsubara modes and the electric scale have been matched
into a three-dimensional static effective theory, the remaining magnetic
Yang--Mills action has the form
\[
  \frac1{4g_3^2}\int d^3x\,\operatorname{tr}(F_{ij}F_{ij}).
\]
In three dimensions \(g_3^2\) has mass dimension one.  A vacuum free-energy
density of the magnetic theory has mass dimension three.  If no scale other
than \(g_3^2\) remains in the confining magnetic sector, dimensional analysis
forces the nonperturbative contribution to be \(c(g_3^2)^3\), where \(c\) is
a dimensionless number defined by the three-dimensional Yang--Mills theory.
The four-dimensional pressure is \(T\) times the static three-dimensional
free-energy density, so the order is \(g^6T^4\).

Perturbation theory cannot determine \(c\), because it expands around a
massless three-dimensional gauge theory whose infrared behavior is precisely
the magnetic theory being asked to generate the scale.  This dimensional
argument does not prove the magnetic mass gap; it identifies the order at which the
nonperturbative magnetic input first enters the high-temperature expansion.
\end{remark}

\section{Transport and Quark--Gluon Plasma Observables}

The quark--gluon plasma is not defined by the presence of free quarks and
gluons.  In this monograph it is the thermal regime of QCD in which the
dominant excitations and response functions are better organized by color
degrees of freedom, screening, and hydrodynamic transport than by a dilute
gas of hadrons.  The precise observables are correlation functions.

The shear viscosity, bulk viscosity, and electrical conductivity are defined
by the Kubo limits of Chapter~\ref{ch:thermal-spectral-kubo-transport}.  For
example, with \(T_{xy}\) a spatial off-diagonal stress tensor component,
\[
  \eta
  =
  \lim_{\omega\downarrow0}
  \frac{1}{2\omega}\,
  \rho_{T_{xy}T_{xy}}(\omega,\mathbf0)
\]
when the limit exists after the thermodynamic limit.  The definition is
nonperturbative; a kinetic or holographic computation is an approximation to
this spectral limit, not the definition of the coefficient.

\begin{controlledapproximation}[QCD hydrodynamic response window]
\label{ca:qcd-hydrodynamic-response-window}
Fix a thermal QCD phase at \((T,\mu_B)\) for which the continuum real-time
retarded kernels of the gauge-invariant stress tensor and global currents
exist.  A claim that QCD is described by first-order hydrodynamics in a
specified channel consists of more than the Kubo limits.  It requires a
compact low-frequency and long-wavelength window \(W_\epsilon\), a subtraction
of declared local contact terms, and a comparison
\begin{equation}
  K^R_{\rm QCD}(\omega,\mathbf k)
  =
  K^R_{\rm hydro}(\omega,\mathbf k;\mathcal T_{\rm QCD})
  +
  P_{\rm loc}(\omega,\mathbf k)
  +
  R_\epsilon(\omega,\mathbf k),
  \label{eq:qcd-hydro-response-comparison}
\end{equation}
where \(\mathcal T_{\rm QCD}\) is the QCD transport datum
\[
  \mathcal T_{\rm QCD}
  =
  \bigl(
    p(T,\mu_B),\ \chi_{ab}(T,\mu_B),\
    \eta(T,\mu_B),\ \zeta(T,\mu_B),\ \Sigma_{ab}(T,\mu_B),\ldots
  \bigr)
\]
extracted from thermodynamic derivatives and Kubo spectral slopes.  Here
\(\chi_{ab}\) and \(\Sigma_{ab}\) are the susceptibility and conductivity
matrices for the conserved charge sector after the stress-tensor variables
have been projected into the chosen hydrodynamic frame.  The remainder
statement must be in the same scaling topology as the hydrodynamic mode:
\(\omega\sim k^2\) for shear and diffusion, and \(\omega\sim k\) for sound.
Thus the shear channel uses
\[
  D_\eta=\frac{\eta}{w},
  \qquad
  \omega_{\rm shear}=-\ii D_\eta k^2+O(k^4),
  \qquad
  w=\epsilon+p,
\]
a decoupled baryon diffusion channel uses
\[
  D_B=\frac{\sigma_B}{\chi_B},
  \qquad
  \omega_{\rm diff}=-\ii D_B k^2+O(k^4),
\]
while at nonzero density or with several conserved charges \(D_B\) is replaced
by the relevant eigenvalue after the sound variables have been separated.  The
same longitudinal matrix also changes the sound attenuation.  In a one-charge
Landau-frame chart write
\begin{equation}
  \beta_1=\left(\frac{\partial p}{\partial\epsilon}\right)_{n_B},
  \qquad
  \beta_2=\left(\frac{\partial p}{\partial n_B}\right)_{\epsilon},
  \qquad
  \alpha_1=T\left(\frac{\partial(\mu_B/T)}{\partial\epsilon}\right)_{n_B},
  \qquad
  \alpha_2=T\left(\frac{\partial(\mu_B/T)}{\partial n_B}\right)_{\epsilon},
  \label{eq:qcd-charged-sound-thermodynamic-derivatives}
\end{equation}
where the dissipative current is
\(J_B^i=n_Bu^i-\Sigma_B^{\rm inc}T\,\partial_i(\mu_B/T)+O(\partial^2)\).
Then the sound channel uses
\[
  \omega_{\rm sound,\pm}
  =
  \pm c_s k
  -\frac{\ii}{2}\Gamma_s k^2
  +O(k^3),
  \qquad
  c_s^2=\beta_1+\frac{n_B}{w}\beta_2,
\]
and
\begin{equation}
  \Gamma_s
  =
  \gamma_{\rm visc}
  +
  \Gamma_{\rm cond},
  \qquad
  \gamma_{\rm visc}
  =
  \frac{\zeta+2\eta(d-1)/d}{w},
  \qquad
  \Gamma_{\rm cond}
  =
  \frac{\Sigma_B^{\rm inc}\beta_2}{c_s^2}
  \left(\alpha_1+\frac{n_B}{w}\alpha_2\right).
  \label{eq:qcd-charged-sound-attenuation}
\end{equation}
For several charges this scalar expression is replaced by the corresponding
charge block of the coupled longitudinal matrix.  The neutral formula is the
special case in which the conductive term vanishes, for example by charge
conjugation at \(\mu_B=0\).
Here \(d\) is the number of spatial dimensions and \(c_s^2\) is computed from
the QCD pressure with the stated thermodynamic constraint.  A bound such as
\begin{equation}
  \sup_{(\omega,\mathbf k)\in W_\epsilon}
  |R_\epsilon(\omega,\mathbf k)|
  \le C\,\epsilon^{N+1}
  \label{eq:qcd-hydro-response-remainder}
\end{equation}
is the additional microscopic input.  It asserts that nonconserved QCD
excitations and long-time tails do not produce extra singular terms in the
window being approximated.  Without a statement of this form, Kubo formulae
define transport coefficients, but they do not by themselves derive
hydrodynamics from QCD.
\end{controlledapproximation}

\begin{controlledapproximation}[Euclidean-to-retarded transport extraction]
\label{ca:qcd-euclidean-retarded-transport-extraction}
The preceding window is a real-time statement.  A lattice calculation, by
contrast, usually supplies Euclidean thermal correlators.  For a
gauge-invariant channel \(X\), after removing declared zero-mode and contact
terms, the positive-frequency spectral representation has the form
\begin{equation}
  G_{E,X}(\tau,\mathbf k)-G_{0,X}-C_{E,X}(\tau,\mathbf k)
  =
  \int_0^\infty \frac{d\omega}{2\pi}\,
  \rho_X(\omega,\mathbf k)K_\beta(\tau,\omega),
  \qquad
  K_\beta(\tau,\omega)
  =
  \frac{\cosh[\omega(\beta/2-\tau)]}{\sinh(\beta\omega/2)} .
  \label{eq:qcd-euclidean-spectral-kernel}
\end{equation}
Equivalently, one may write the same kernel with
\(\widehat\rho_X:=\rho_X/(2\pi)\) and the measure \(d\omega\).  The
normalization is part of the inverse problem: omitting the \(1/(2\pi)\) factor
rescales every matrix element that connects Euclidean data to the real-time
Kubo slope.
The transport coordinate in this channel is a low-frequency real-time slope,
for example
\[
  \lambda_X
  =
  \lim_{\omega\downarrow0}
  \frac{\rho_X(\omega,\mathbf0)}{2\omega}
\]
with the appropriate stress-tensor, bulk, or current source normalization.
Therefore a number extracted from \(G_{E,X}(\tau)\) is not yet a QCD
viscosity or conductivity.  The extraction datum must include the Euclidean
linear functionals being used, the contact and zero-mode subtraction, the
ultraviolet tail or OPE input, the low-frequency spectral class, and a
controlled inverse map to the slope \(\lambda_X\).

One convenient finite-window notation is the following.  Let \(L_i\) be the
time smearing or Matsubara functionals retained from the Euclidean
calculation and let \(a_\alpha\) be the low-frequency spectral coordinates
of a declared ansatz or variational class.  After contact and tail
subtraction,
\begin{equation}
  g_i
  =
  L_i(G_{E,X})-L_i(C_{E,X})-L_i(G_{E,X}^{\rm tail})
  =
  \sum_\alpha M_{i\alpha}a_\alpha+r_i,
  \qquad
  \lambda_X=\ell_X(a).
  \label{eq:qcd-euclidean-inverse-transport-map}
\end{equation}
Here \(M_{i\alpha}\) is constructed using the same normalized measure as
\eqref{eq:qcd-euclidean-spectral-kernel}.  If the coordinates \(a_\alpha\)
parametrize \(\rho_X\) rather than \(\widehat\rho_X\), the factor
\(1/(2\pi)\) is included in \(M_{i\alpha}\); if they parametrize
\(\widehat\rho_X\), it is absorbed into \(a_\alpha\).  The stability norm below
must be computed in that same coordinate system.
An accepted transport extraction supplies a reconstruction map
\(\widehat a=\mathcal I_X(g)\), a stability norm, and a residual estimate
\begin{equation}
  |\widehat\lambda_X-\lambda_X|
  \le
  \|\ell_X D\mathcal I_X\|\,
  \bigl(
    \|R_E\|+\|R_{\rm contact}\|+\|R_{\rm tail}\|
  \bigr)
  +
  R_{\rm ansatz}+R_{\rm FV}+R_{\rm cont}+R_{\rm AC}.
  \label{eq:qcd-euclidean-retarded-transport-budget}
\end{equation}
Here \(R_{\rm AC}\) records the analytic-continuation or spectral-inference
instability, \(R_{\rm FV}\) the finite-volume distortion of small
frequencies, and \(R_{\rm cont}\) the continuum and operator-renormalization
limit.  The ansatz residual is physical: it is the assertion that the
hydrodynamic peak, nearby nonhydrodynamic modes, and continuum tail have been
separated in the chosen channel.  Positivity, oddness of the commutator
spectral density, and KMS are constraints on this inverse problem; they are
not an inverse problem by themselves.

This distinction is especially important for QCD transport.  A Euclidean
midpoint value, a small collection of Matsubara points, or a finite-\(k\)
Euclidean pole fit can be compatible with several low-frequency spectral
slopes unless the spectral class and stability estimate are stated.
Conversely, local metric or current counterterms can change Euclidean
moments without changing the commutator spectral slope.  The shear, bulk,
and charge windows below are therefore real-time spectral extraction windows.  A
Euclidean computation enters them only after it has passed
\eqref{eq:qcd-euclidean-inverse-transport-map} and
\eqref{eq:qcd-euclidean-retarded-transport-budget}; otherwise it is evidence
about an imaginary-time correlator, not yet a transport coefficient of QCD.
\end{controlledapproximation}

\begin{controlledapproximation}[Finite shear spectral window]
\label{ca:qcd-finite-shear-spectral-window}
The response-window statement can be tested more directly in a finite
frequency interval.  Let
\(\pi_x=T^{0x}\) be a transverse momentum density and take momentum
\(\mathbf k=k\hat y\).  If the shear pole is isolated in the QCD retarded
kernel, the contact-subtracted pole part may be represented, in the
\(\rho=-2\operatorname{Im}G^R\) convention of
Chapter~\ref{ch:thermal-spectral-kubo-transport}, by
\begin{equation}
  G_{\pi_x\pi_x}^{R,{\rm pole}}(\omega,k)
  =
  -\frac{w\,\gamma_k}{\gamma_k-\ii\omega}.
  \label{eq:qcd-shear-retarded-pole-kernel}
\end{equation}
Its pole lies at \(\omega=-\ii\gamma_k\).  The corresponding
positive-frequency spectral density is
\begin{equation}
  \rho^{\rm hyd}_{\pi_x\pi_x}(\omega,k)
  =
  \frac{2w\,\gamma_k\,\omega}{\omega^2+\gamma_k^2},
  \qquad
  \gamma_k=D_\eta k^2=\frac{\eta}{w}k^2,
  \label{eq:qcd-shear-hydro-spectral-peak}
\end{equation}
up to the contact subtraction and the same frame choice used in
\eqref{eq:qcd-hydro-response-comparison}.
Thus an extraction can be checked at the response-function level: the width is
the pole location, the residue is the static momentum susceptibility \(w\), and
the viscosity is reconstructed as \(w\gamma_k/k^2\).  A finite spectral
extraction is
therefore not the statement ``read off \(\eta\) from a plotted peak.''  It
requires scales
\begin{equation}
  \gamma_k\ll\Omega\ll \tau_{\rm micro}^{-1},
  \label{eq:qcd-shear-spectral-window-scale-separation}
\end{equation}
a subtracted spectral density \(\rho^{\rm sub}\), and a bound
\begin{equation}
  \rho^{\rm sub}_{\pi_x\pi_x}(\omega,k)
  =
  \rho^{\rm hyd}_{\pi_x\pi_x}(\omega,k)
  +
  \rho^{\rm reg}_{\pi_x\pi_x}(\omega,k)
  +
  \rho^{\rm near}_{\pi_x\pi_x}(\omega,k)
  +
  R_{\Omega,k}(\omega).
  \label{eq:qcd-shear-spectral-window-decomposition}
\end{equation}
Here \(\rho^{\rm reg}\) is the analytic low-frequency background from gapped
nonhydrodynamic QCD excitations, \(\rho^{\rm near}\) is any additional
near-critical or long-time-tail contribution that is singular in the same
window, and \(R_{\Omega,k}\) is the finite-regulator and continuum-window
remainder.  The peak residue is checked by
\begin{equation}
  \mathcal A_\Omega(k)
  =
  \frac1\pi\int_0^\Omega
  \frac{\rho^{\rm sub}_{\pi_x\pi_x}(\omega,k)}{\omega}\,d\omega
  =
  \frac{2w}{\pi}\arctan\!\left(\frac{\Omega}{\gamma_k}\right)
  +
  R_A(\Omega,k),
  \label{eq:qcd-shear-spectral-window-area}
\end{equation}
so that
\begin{equation}
  \left|\mathcal A_\Omega(k)-w\right|
  \le
  \frac{2w}{\pi}\frac{\gamma_k}{\Omega}
  +
  \frac1\pi\int_0^\Omega
  \frac{|\rho^{\rm reg}|+|\rho^{\rm near}|+|R_{\Omega,k}|}{\omega}\,d\omega .
  \label{eq:qcd-shear-window-residue-bound}
\end{equation}
If the width \(\widehat\gamma_k\) and residue \(\widehat w\) have been
isolated with errors \(\delta\gamma_k\) and \(\delta w\), the corresponding
transport estimate obeys
\begin{equation}
  \left|
  \frac{\widehat w\,\widehat\gamma_k}{k^2}-\eta
  \right|
  \le
  \frac{
    w|\delta\gamma_k|
    +
    \gamma_k|\delta w|
    +
    |\delta w|\,|\delta\gamma_k|
  }{k^2}
  +
  R_{\rm model}.
  \label{eq:qcd-shear-window-eta-error-budget}
\end{equation}
This is the finite-window content behind a QCD shear-viscosity extraction.  The
Kubo slope at \(k=0\), the finite-\(k\) hydrodynamic peak width, the enthalpy
residue \(w=\epsilon+p\), the entropy density \(s=\partial_Tp\), and any
kinetic or holographic relaxation scale are different pieces of data.  A
comparison such as \(\eta/s\) is meaningful only after these data have been
defined in the same QCD phase specification and after the regular background,
near-critical modes, long-time tails, Euclidean reconstruction prior, and
continuum/finite-volume errors have been included in the window budget.
\end{controlledapproximation}

\begin{controlledapproximation}[Finite bulk and sound spectral window]
\label{ca:qcd-finite-bulk-sound-spectral-window}
The scalar stress channel is subtler than the shear channel because the trace
of the spatial stress is not itself the dissipative bulk-pressure operator.
At fixed QCD thermodynamic constraint, write the fluctuation
\(\delta T^{00}=T^{00}-\langle T^{00}\rangle\) and subtract the equilibrium
pressure trace by
\[
  \delta P_{\rm tr}
  =
  \frac1d\bigl(T^i{}_i-\langle T^i{}_i\rangle\bigr).
\]
The bulk source which is orthogonal to the conserved energy density is
\begin{equation}
  \mathcal B
  =
  \delta P_{\rm tr}-c_s^2\,\delta T^{00},
  \label{eq:qcd-bulk-subtracted-source}
\end{equation}
with the same contact-term convention and hydrodynamic frame as in
\eqref{eq:qcd-hydro-response-comparison}.  The bulk-viscosity Kubo datum is
therefore
\begin{equation}
  \zeta
  =
  \lim_{\omega\downarrow0}
  \frac{1}{2\omega}\,
  \rho_{\mathcal B\mathcal B}(\omega,\mathbf0),
  \label{eq:qcd-bulk-subtracted-kubo}
\end{equation}
not the raw slope of the trace--trace correlator before the pressure,
energy-density, and contact subtractions have been fixed.  Equivalently, in a
finite spectral reconstruction the low-frequency scalar slope must be formed
from the subtracted matrix
\begin{equation}
  S_{\mathcal B}
  =
  S_{PP}
  -2c_s^2S_{PE}
  +c_s^4S_{EE},
  \qquad
  S_{XY}
  =
  \lim_{\omega\downarrow0}
  \frac{\rho_{XY}^{\rm sub}(\omega,\mathbf0)}{2\omega},
  \label{eq:qcd-bulk-slope-subtraction}
\end{equation}
after the thermodynamic derivative used to define \(c_s^2\) has been stated.

The same coefficient can be reached from the longitudinal sound pole only
after the shear and conductive contributions have also been subtracted.  In
the one-charge notation of
\eqref{eq:qcd-charged-sound-thermodynamic-derivatives}, the linearized
Landau-frame longitudinal equations for momentum \(k\hat x\) are
\begin{align}
  \partial_t\delta\epsilon+\ii k\,\pi_\parallel&=0,
  \nonumber\\
  \partial_t\pi_\parallel
  +\ii k\,\beta_1\delta\epsilon
  +\ii k\,\beta_2\delta n_B
  +\gamma_{\rm visc}k^2\pi_\parallel&=0,
  \nonumber\\
  \partial_t\delta n_B
  +\ii k\,\frac{n_B}{w}\pi_\parallel
  +\Sigma_B^{\rm inc}k^2(\alpha_1\delta\epsilon+\alpha_2\delta n_B)&=0.
  \label{eq:qcd-charged-sound-linearized-matrix}
\end{align}
Their characteristic polynomial has the small-\(k\) roots
\begin{equation}
  \omega_+(k)
  =
  c_s k-\frac{\ii}{2}\Gamma_s k^2+O(k^3),
  \qquad
  \Gamma_s(k)=-\frac{2\,\operatorname{Im}\omega_+(k)}{k^2},
  \label{eq:qcd-sound-width-estimator}
\end{equation}
with \(\Gamma_s\) as in \eqref{eq:qcd-charged-sound-attenuation}.  Therefore
the finite-\(k\) bulk estimate at finite density is
\begin{equation}
  \widehat\zeta_k
  =
  \widehat w\,
  \bigl(\widehat\Gamma_s(k)-\widehat\Gamma_{\rm cond}\bigr)
  -
  \frac{2(d-1)}{d}\,\widehat\eta .
  \label{eq:qcd-bulk-from-sound-estimator}
\end{equation}
Thus a plotted sound attenuation width is not a bulk-viscosity measurement by
itself.  It becomes one only after the enthalpy \(w\), the shear viscosity
\(\eta\), the sound-speed thermodynamic constraint, the charge-diffusion
projection at finite density, the conductive attenuation
\(\Gamma_{\rm cond}\), and the \(O(k^3)\) dispersion remainder have all been
supplied in the same QCD phase specification.  At \(\mu_B=0\), or at a special
decoupling point with \(\Gamma_{\rm cond}=0\),
\eqref{eq:qcd-bulk-from-sound-estimator} reduces to the neutral-fluid
estimator.  A response-window
error budget has the form
\begin{align}
  |\widehat\zeta_k-\zeta|
  &\le
  w\,(\delta\Gamma_s+\delta\Gamma_{\rm cond})
  +(\Gamma_s-\Gamma_{\rm cond})\,\delta w
  +\delta w\,(\delta\Gamma_s+\delta\Gamma_{\rm cond})
  +\frac{2(d-1)}d\,\delta\eta
  \nonumber\\
  &\quad
  +R_{k^3}
  +R_{\rm reg}
  +R_{\rm crit}
  +R_{\rm cont}
  +R_{\rm therm},
  \label{eq:qcd-bulk-sound-error-budget}
\end{align}
where \(\delta\Gamma_{\rm cond}\) includes the conductivity and thermodynamic
derivative errors in \(\Sigma_B^{\rm inc}\), \(\alpha_{1,2}\), \(\beta_2\),
\(n_B/w\), and \(c_s^2\).  Here \(R_{k^3}\) is the finite-\(k\) dispersion
error, \(R_{\rm reg}\) the analytic nonhydrodynamic background,
\(R_{\rm crit}\) any extra scalar near-critical peak or cut, \(R_{\rm cont}\)
the continuum and finite-volume subtraction remainder, and \(R_{\rm therm}\)
the pressure-derivative error in \(c_s^2\), \(w\), and the bulk source.  The
near-critical term is displayed separately because the QCD scalar channel is
precisely where critical fluctuations can collide with the hydrodynamic sound
window; such weight is not a harmless regular background.
\end{controlledapproximation}

\begin{proposition}[Momentum-projected baryon diffusion current]
\label{prop:qcd-incoherent-baryon-current}
In a translationally invariant QCD phase at nonzero baryon density, the raw
baryon current is not itself the diffusive current.  Let
\[
  n_B=\frac{\partial p}{\partial\mu_B},\qquad
  w=\epsilon+p,
  \qquad P^i=T^{0i},
\]
and let \(J_B^i\) be the spatial baryon-number current.  In a homogeneous
fluid state boosted by a small velocity \(\delta u^i\),
\[
  \delta\langle P^i\rangle=w\,\delta u^i,\qquad
  \delta\langle J_B^i\rangle=n_B\,\delta u^i .
\]
Thus the static susceptibility Gram matrix in the vector sector contains
\[
  \chi_{P^iP^j}=w\,\delta^{ij},
  \qquad
  \chi_{J_B^iP^j}=n_B\,\delta^{ij}.
\]
The momentum-orthogonal baryon current is therefore
\begin{equation}
  J_{\rm inc}^i
  =
  J_B^i-\frac{n_B}{w}T^{0i}.
  \label{eq:qcd-incoherent-baryon-current}
\end{equation}
Writing \(\chi_{J_B^iJ_B^j}=\chi_{JJ}\delta^{ij}\), it obeys
\begin{equation}
  \chi_{J_{\rm inc}^iP^j}=0,\qquad
  \chi_{J_{\rm inc}^iJ_{\rm inc}^j}
  =
  \left(\chi_{JJ}-\frac{n_B^2}{w}\right)\delta^{ij}.
  \label{eq:qcd-incoherent-current-gram}
\end{equation}
Consequently the baryon conductivity entering the diffusive part of
\(\mathcal T_{\rm QCD}\) is the regular Kubo slope of
\(J_{\rm inc}\),
\begin{equation}
  \Sigma_B^{\rm inc}
  =
  \lim_{\omega\downarrow0}
  \frac1{2\omega}
  \rho^{\rm reg}_{J_{\rm inc}^iJ_{\rm inc}^i}(\omega,\mathbf0),
  \label{eq:qcd-incoherent-baryon-conductivity}
\end{equation}
with the Drude sector separated as in
\eqref{eq:kubo-drude-regular-decomposition}.  At \(n_B=0\),
\(J_{\rm inc}=J_B\).  At \(n_B\ne0\), the raw
\(J_BJ_B\) response contains a convective contribution from the exactly
conserved momentum.  If translations are weakly broken this contribution
becomes a narrow momentum-relaxation peak; that peak is a property of the
relaxation mechanism, not the intrinsic QCD baryon-diffusion coefficient.
\end{proposition}

\begin{proof}
The ideal constitutive relations in the local rest frame are
\[
  T^{0i}=w\,u^i+O(\partial),\qquad
  J_B^i=n_B\,u^i+O(\partial).
\]
Differentiating the expectation values with respect to a uniform velocity
source gives the displayed static overlaps.  Equation
\eqref{eq:qcd-incoherent-baryon-current} is then ordinary Gram--Schmidt
projection in the static susceptibility inner product, so
\[
  \chi_{J_{\rm inc}P}
  =
  \chi_{J_BP}-\frac{n_B}{w}\chi_{PP}
  =
  n_B-\frac{n_B}{w}w
  =
  0.
\]
The second identity in \eqref{eq:qcd-incoherent-current-gram} is the same
projection applied to the norm of the current.  Positivity is the Schur
complement positivity of the static susceptibility matrix.  The last
statement is then the Drude separation of
\eqref{eq:kubo-drude-regular-decomposition}: any current with a nonzero
static overlap with a conserved momentum has a singular zero-frequency
sector, while the projected current is the object whose regular spectral
slope can be matched to the diffusive hydrodynamic coefficient.
\end{proof}

\begin{controlledapproximation}[Finite charge diffusion spectral window]
\label{ca:qcd-finite-charge-diffusion-spectral-window}
After the convective current in
Proposition~\ref{prop:qcd-incoherent-baryon-current} has been removed, a
conserved charge conductivity can be tested in a finite density-response
window.  Work first in one momentum-orthogonal charge eigenchannel.  Let
\(\chi_q>0\) be the static susceptibility, \(\Sigma_q^{\rm inc}\) the
intrinsic conductivity, and
\[
  D_q=\frac{\Sigma_q^{\rm inc}}{\chi_q},
  \qquad
  \gamma_q(k)=D_q k^2 .
\]
For a contact-subtracted density source with momentum
\(\mathbf k=k\hat x\), an isolated diffusion pole has the retarded kernel
\begin{equation}
  G_{n_qn_q}^{R,{\rm pole}}(\omega,k)
  =
  -\frac{\chi_q\,\gamma_q(k)}{\gamma_q(k)-\ii\omega}.
  \label{eq:qcd-charge-diffusion-retarded-pole}
\end{equation}
In the same \(\rho=-2\operatorname{Im}G^R\) convention as the shear window,
\begin{equation}
  \rho^{\rm hyd}_{n_qn_q}(\omega,k)
  =
  \frac{2\chi_q\,\gamma_q(k)\,\omega}
       {\omega^2+\gamma_q(k)^2}.
  \label{eq:qcd-charge-diffusion-spectral-peak}
\end{equation}
Thus the finite-\(k\) width determines \(D_q\), while the residue determines
\(\chi_q\).  The conductivity is their product,
\[
  \Sigma_q^{\rm inc}=\chi_q\,\frac{\gamma_q(k)}{k^2},
\]
not the width by itself and not the raw current Kubo slope before momentum
projection.

A controlled finite-window extraction requires
\begin{equation}
  \gamma_q(k)\ll\Omega\ll \tau_{\rm micro}^{-1}
  \label{eq:qcd-charge-diffusion-window-scale-separation}
\end{equation}
and a decomposition
\begin{equation}
  \rho^{\rm sub}_{n_qn_q}(\omega,k)
  =
  \rho^{\rm hyd}_{n_qn_q}(\omega,k)
  +
  \rho^{\rm reg}_{n_qn_q}(\omega,k)
  +
  \rho^{\rm near}_{n_qn_q}(\omega,k)
  +
  R^{(q)}_{\Omega,k}(\omega).
  \label{eq:qcd-charge-diffusion-window-decomposition}
\end{equation}
Here \(\rho^{\rm reg}\) is analytic low-frequency weight,
\(\rho^{\rm near}\) is any critical or long-time-tail charge mode in the same
scaling window, and \(R^{(q)}_{\Omega,k}\) is the finite-volume,
continuation, and continuum-window remainder.  The susceptibility residue is
checked by
\begin{equation}
  \mathcal C_\Omega(k)
  =
  \frac1\pi\int_0^\Omega
  \frac{\rho^{\rm sub}_{n_qn_q}(\omega,k)}{\omega}\,d\omega
  =
  \frac{2\chi_q}{\pi}
  \arctan\!\left(\frac{\Omega}{\gamma_q(k)}\right)
  +
  R_C(\Omega,k),
  \label{eq:qcd-charge-diffusion-window-area}
\end{equation}
so
\begin{equation}
  \left|\mathcal C_\Omega(k)-\chi_q\right|
  \le
  \frac{2\chi_q}{\pi}\frac{\gamma_q(k)}{\Omega}
  +
  \frac1\pi\int_0^\Omega
  \frac{|\rho^{\rm reg}|+|\rho^{\rm near}|+|R^{(q)}_{\Omega,k}|}{\omega}\,
  d\omega .
  \label{eq:qcd-charge-diffusion-residue-bound}
\end{equation}
If the fitted width and residue are
\(\widehat\gamma_q=\gamma_q+\delta\gamma_q\) and
\(\widehat\chi_q=\chi_q+\delta\chi_q\), then
\begin{equation}
  \left|
  \frac{\widehat\chi_q\,\widehat\gamma_q}{k^2}
  -
  \Sigma_q^{\rm inc}
  \right|
  \le
  \frac{
    \chi_q|\delta\gamma_q|
    +
    \gamma_q|\delta\chi_q|
    +
    |\delta\chi_q|\,|\delta\gamma_q|
  }{k^2}
  +
  R_{\rm charge}.
  \label{eq:qcd-charge-diffusion-conductivity-budget}
\end{equation}
For several exactly conserved charges, the same statement is applied after
diagonalizing the momentum-orthogonal matrix
\(\Sigma^{\rm inc}\chi^{-1}\) in the charge sector and after separating the
sound variables at finite density.  Otherwise a longitudinal density peak can
mix with sound, while a raw current correlator can contain the convective
Drude contribution of the conserved momentum.  Neither effect is an intrinsic
QCD charge-conductivity datum.
\end{controlledapproximation}

\begin{controlledapproximation}[QCD transport closure from finite windows]
\label{ca:qcd-transport-closure-window}
The three preceding finite-window statements become a hydrodynamic
description of QCD only after their outputs are assembled into one
same-state transport datum.  Fix the phase, regulator trajectory,
renormalized stress tensor and currents, thermodynamic constraint, and
hydrodynamic frame.  The first-order datum is
\begin{equation}
  \mathfrak T_{\rm QCD}^{(1)}
  =
  \bigl(
    w,\ n_B,\ c_s^2,\ \beta_{1,2},\ \alpha_{1,2},\
    \eta,\ \zeta,\ \chi^{\perp}_{ab},\ \Sigma_{ab}^{\rm inc}
  \bigr)_{(T,\mu_B)},
  \qquad
  w=\epsilon+p ,
  \label{eq:qcd-first-order-transport-datum}
\end{equation}
where \(\chi^{\perp}\) and \(\Sigma^{\rm inc}\) are evaluated after the
momentum and sound variables have been projected out of the charge sector.
This datum is not the disjoint list of best available numbers in several
channels.  The same \(w\), \(c_s^2\), contact convention, and frame must be
used in the shear peak, the scalar sound reconstruction, and the projected
charge diffusion problem.

For a single retained baryon charge, the closure conditions are
\begin{equation}
  D_\eta=\frac{\eta}{w},
  \qquad
  \Gamma_s
  =
  \frac{\zeta+2(d-1)\eta/d}{w}
  +
  \frac{\Sigma_B^{\rm inc}\beta_2}{c_s^2}
  \left(\alpha_1+\frac{n_B}{w}\alpha_2\right),
  \qquad
  D_B
  =
  \frac{\Sigma_B^{\rm inc}(\alpha_2\beta_1-\alpha_1\beta_2)}
       {c_s^2},
  \label{eq:qcd-transport-closure-diffusivities}
\end{equation}
with the neutral formula recovered only when the second term in \(\Gamma_s\)
vanishes.  With several conserved charges these scalar expressions are
replaced by the sound and diffusion eigenvalues of the coupled longitudinal
susceptibility-conductivity matrix.  The pole statements in a common
low-energy window are
\begin{align}
  \omega_{\rm shear}
  &=
  -\ii D_\eta k^2+O(k^4),
  \nonumber\\
  \omega_{\rm sound,\pm}
  &=
  \pm c_s k-\frac{\ii}{2}\Gamma_s k^2+O(k^3),
  \nonumber\\
  \omega_{\rm diff,B}
  &=
  -\ii D_B k^2+O(k^4).
  \label{eq:qcd-transport-closure-poles}
\end{align}
The finite-window comparison must therefore carry a combined residual
\begin{equation}
  R_{\rm hydro}
  =
  R_{\rm shear}
  +R_{\rm bulk}
  +R_{\rm charge}
  +R_{\rm cond}
  +R_{\rm therm}
  +R_{\rm frame}
  +R_{\rm cross}
  +R_{\rm cont}.
  \label{eq:qcd-transport-closure-residual}
\end{equation}
Here \(R_{\rm cond}\) records the uncertainty in the finite-density conductive
sound term, including the conductivity matrix and the thermodynamic derivative
combination in \eqref{eq:qcd-charged-sound-attenuation}.  The term
\(R_{\rm cross}\) records the agreement of the thermodynamic state and
subtractions across channels, while \(R_{\rm frame}\) records the change from
the microscopic current basis to the chosen Landau-frame or equivalent
hydrodynamic variables.  A shear extraction plus a pressure curve does not
determine bulk viscosity; a sound attenuation width gives \(\zeta\) only
after the shear and conductive subtractions in
\eqref{eq:qcd-bulk-from-sound-estimator}; and a charge diffusion width gives
\(\Sigma_q^{\rm inc}\) only after the susceptibility residue and
momentum-projected current have been supplied.  Likewise, transport
coefficients extracted at nearby but distinct \((T,\mu_B)\), with different
renormalized stress tensors, or with different frame conventions do not form
\(\mathfrak T_{\rm QCD}^{(1)}\).

This closure statement is still conditional on the microscopic response
window of \eqref{eq:qcd-hydro-response-comparison}.  Its role is to turn the
separate spectral extractions into a physical hydrodynamic amplitude
prediction for QCD correlators in the specified window, and to make explicit
which extra information is required beyond the individual Kubo limits.
\end{controlledapproximation}

\begin{remark}[The \(\eta/s\) lower-bound claim]
The value \(1/(4\pi)\) is an important result in a class of strongly coupled
large-\(N\) theories with classical gravity duals.  It is not a theorem of
QCD.  When this number is used as a comparison scale in QCD phenomenology,
the statement is a quantitative comparison of transport data or models, not
a rigorous lower bound derived from the QCD axioms.
\end{remark}

\section{Finite Baryon Density and the Sign Problem}

At real \(\mu_B\), the Euclidean Dirac operator is no longer anti-Hermitian
up to the mass term.  In continuum notation the temporal derivative is
shifted by \(-\mu_q\), and on the lattice forward and backward temporal hops
are weighted asymmetrically.  The determinant need not be real and
nonnegative.

At \(\mu_B=0\), a vectorlike pair of equal-mass quark flavors gives a
nonnegative Euclidean determinant.  At real \(\mu_B\ne0\), the same
positivity argument fails because the Dirac operator loses the
\(\gamma_5\)-Hermiticity relation that pairs eigenvalues into complex
conjugates at fixed gauge field.
At \(\mu_B=0\), the lattice or continuum Euclidean Dirac operator obeys
\[
  \gamma_5(D_E+m)\gamma_5=(D_E+m)^\dagger
\]
for real mass \(m\).  Therefore
\(\det(D_E+m)\) is real, and for two degenerate flavors the product is
\(|\det(D_E+m)|^2\ge0\).  This is the measure-positivity input behind
importance sampling.

A real quark chemical potential changes the temporal part of the operator by
a term that is Hermitian rather than anti-Hermitian in Euclidean signature.
Equivalently, the lattice temporal hopping factors \(e^{a\mu_q}\) and
\(e^{-a\mu_q}\) are not related by the same unitary conjugation that holds at
\(\mu_q=0\).  The displayed \(\gamma_5\)-Hermiticity relation is replaced by
a relation connecting \(\mu_q\) to \(-\mu_q\).  At fixed real \(\mu_q\) the
determinant is generally complex, so it cannot be used directly as a
positive probability weight.

\subsection{Imaginary Chemical Potential and Roberge--Weiss Periodicity}
\label{subsec:qcd-imaginary-mu-roberge-weiss}

Imaginary chemical potential is a different, exactly defined thermal boundary
condition from real finite density.  It is useful because it restores
positivity in vectorlike pairs and has an exact center
periodicity at finite regulator.

\begin{definition}[Imaginary quark and baryon chemical-potential angles]
\label{def:qcd-imaginary-mu-angles}
Let
\[
  \mu_q=iT\theta_q,\qquad
  \mu_B=iT\theta_B,\qquad
  \theta_B=N_c\theta_q .
\]
Equivalently, the fundamental quark field obeys the Euclidean thermal
boundary condition
\[
  q(\beta,\mathbf x)=-e^{i\theta_q}q(0,\mathbf x).
\]
For a finite lattice regulator, write the corresponding partition function
as \(Z_{a,V}(\theta_q)\).  The regulator is assumed to use compact
\(SU(N_c)\) links, a gauge action invariant under center-twisted temporal
gauge transformations, and a gauge-covariant Dirac operator whose imaginary
chemical potential is this boundary twist.
\end{definition}

\begin{lemma}[Finite-regulator Roberge--Weiss periodicity]
\label{lem:qcd-roberge-weiss-periodicity}
At finite regulator and finite spatial volume,
\[
  Z_{a,V}\!\left(\theta_q+\frac{2\pi k}{N_c}\right)
  =
  Z_{a,V}(\theta_q),
  \qquad k\in\mathbb Z .
\]
Equivalently, as a function of the baryon angle,
\[
  Z_{a,V}(\theta_B+2\pi k)=Z_{a,V}(\theta_B).
\]
For a vectorlike pair of equal-mass flavors, the Euclidean determinant at
real \(\theta_q\) is nonnegative.  Therefore imaginary-\(\mu_B\) simulations
probe a positive finite-regulator measure, while analytic continuation to
real \(\mu_B\) remains a separate limiting and analytic-control problem.
\end{lemma}

\begin{proof}
Let \(z_k=e^{2\pi i k/N_c}\in Z(SU(N_c))\).  A center-twisted temporal gauge
transformation is a gauge transformation \(g(\tau,\mathbf x)\) satisfying
\[
  g(\beta,\mathbf x)=z_k\,g(0,\mathbf x).
\]
Gauge fields in the adjoint representation are periodic under this
transformation, because the center acts trivially on adjoint variables.  The
compact-link Haar measure and gauge action are invariant.

For a fundamental quark field with
\(q(\beta,\mathbf x)=-e^{i\theta_q}q(0,\mathbf x)\), the transformed field
has boundary condition
\[
  q'(\beta,\mathbf x)
  =
  g(\beta,\mathbf x)q(\beta,\mathbf x)
  =
  -e^{i\theta_q}z_k\,g(0,\mathbf x)q(0,\mathbf x)
  =
  -e^{i(\theta_q+2\pi k/N_c)}q'(0,\mathbf x).
\]
Thus the change of integration variables by the center-twisted gauge
transformation maps the finite-regulator path integral at \(\theta_q\) to
the one at \(\theta_q+2\pi k/N_c\).  Since the transformation is invertible
and leaves the measure and gauge action invariant, the two partition
functions are equal.  Multiplying the quark-angle period by \(N_c\) gives
the baryon-angle period \(2\pi\).

For \(\mu_q=iT\theta_q\), the Euclidean chemical-potential term is
anti-Hermitian in the same sense as the ordinary Euclidean Dirac operator.
The massive Dirac operator obeys the \(\gamma_5\)-Hermiticity relation at
fixed gauge field,
\[
  \gamma_5D_{\theta_q,m}\gamma_5=D_{\theta_q,m}^{\dagger}.
\]
Hence its determinant is real, and a pair of equal-mass Dirac flavors gives
\(|\det D_{\theta_q,m}|^2\ge0\).  This proves the positivity statement for
vectorlike pairs.  None of these equalities implies analyticity at real
\(\mu_B\); analytic continuation is controlled only inside zero-free
domains of the complex fugacity or chemical-potential plane.
\end{proof}

\begin{remark}[Roberge--Weiss lines]
\label{rem:qcd-roberge-weiss-lines}
The finite-volume function \(Z_{a,V}(\theta_q)\) is analytic in the imaginary
fugacity.  A Roberge--Weiss transition is therefore a thermodynamic-limit
phenomenon: in high-temperature \(SU(N_c)\) QCD, different center-related
holonomy sectors may exchange dominance at
\[
  \theta_q=\frac{(2r+1)\pi}{N_c},\qquad r\in\mathbb Z .
\]
At finite volume this appears as Lee--Yang zero organization, not as a
literal singularity of \(Z_{a,V}\).  The statement is about the imaginary
chemical-potential problem and does not by itself locate a real-\(\mu_B\)
critical endpoint.
\end{remark}

Consequently, the conjectural critical endpoint and the first-order line in
the \((T,\mu_B)\) plane are not presently consequences of direct
first-principles continuum control at large real \(\mu_B\).  Taylor
expansions around \(\mu_B=0\), imaginary-\(\mu_B\) continuation, reweighting,
and effective models are valuable but must carry their own domains of
validity.

\section{Baryon Susceptibilities and Critical-Point Diagnostics}
\label{sec:qcd-baryon-susceptibilities-criticality}

At \(\mu_B=0\) the Euclidean measure of vectorlike QCD is positive in the
standard lattice discretizations, so derivatives with respect to
\(\mu_B\) give first-principles thermal observables even though direct
importance sampling at real \(\mu_B\ne0\) is obstructed.  The natural
dimensionless source is
\[
  x=\frac{\mu_B}{T}=\beta\mu_B .
\]
At finite regulator and finite volume,
\begin{equation}
  Z_{a,V}(x)
  =
  \operatorname{Tr}_{\Hilb_{a,V}}
  \exp\{-\beta H_{a,V}+xB_{a,V}\}.
  \label{eq:qcd-baryon-finite-x-partition}
\end{equation}
Here \(B_{a,V}\) is the finite-regulator baryon charge, normalized so that a
baryon has charge \(1\), and it is assumed to be an exactly conserved
self-adjoint charge of the regulated theory:
\[
  [H_{a,V},B_{a,V}]=0 .
\]

\begin{definition}[Baryon susceptibilities]
\label{def:qcd-baryon-susceptibilities}
The finite-regulator baryon cumulants and susceptibilities at \(x=0\) are
\begin{equation}
  \kappa_{n,a,V}^{B}(T)
  =
  \left.
  \frac{\partial^n}{\partial x^n}\log Z_{a,V}(x)
  \right|_{x=0},
  \qquad
  \chi_{n,a,V}^{B}(T)
  =
  \frac{1}{VT^3}\kappa_{n,a,V}^{B}(T).
  \label{eq:qcd-baryon-susceptibility-definition}
\end{equation}
The continuum thermodynamic susceptibility \(\chi_n^B(T)\) is the limit of
\(\chi_{n,a,V}^{B}(T)\) along the chosen continuum trajectory, when the
limit exists.
\end{definition}

\paragraph{Baryon susceptibilities as baryon-number cumulants.}
At finite regulator and finite volume,
\begin{equation}
  \chi_{1,a,V}^B
  =
  \frac{\langle B\rangle}{VT^3},
  \qquad
  \chi_{2,a,V}^B
  =
  \frac{\langle B^2\rangle-\langle B\rangle^2}{VT^3},
  \label{eq:qcd-baryon-first-second-cumulants}
\end{equation}
and
\begin{align}
  \chi_{4,a,V}^B
  &=
  \frac{1}{VT^3}
  \bigl(
  \langle B^4\rangle
  -4\langle B^3\rangle\langle B\rangle
  -3\langle B^2\rangle^2
  \notag\\
  &\hspace{3.5cm}
  +12\langle B^2\rangle\langle B\rangle^2
  -6\langle B\rangle^4
  \bigr).
  \label{eq:qcd-baryon-fourth-cumulant}
\end{align}
If charge conjugation is an exact symmetry at \(x=0\) and sends
\(B\mapsto -B\), then all odd finite-volume cumulants vanish.

Because \(B_{a,V}\) is conserved, \(H_{a,V}\) and \(B_{a,V}\) can be
simultaneously decomposed by the spectral theorem, and differentiating
\eqref{eq:qcd-baryon-finite-x-partition} with respect to \(x\) is the
ordinary differentiation of the finite-regulator moment generating function
of the baryon charge in the thermal state.  For any observable \(Y\) commuting
with \(B\), in particular for \(Y=B^k\), one has
\begin{equation}
  \frac{\partial}{\partial x}\langle Y\rangle_x
  =
  \langle YB\rangle_x-\langle Y\rangle_x\langle B\rangle_x .
  \label{eq:qcd-baryon-source-derivative-expectation}
\end{equation}
Applying this identity to \(Y=1,B,B^2,B^3\), or equivalently differentiating
\(\log Z(x)\), gives the displayed cumulants.  If charge conjugation is an
exact symmetry at \(x=0\), the spectral measure of \(B\) in the thermal state
is invariant under \(B\mapsto -B\).  The moment generating function then
satisfies
\[
  Z(x)=Z(-x),
\]
so \(\log Z(x)\) is even near the origin and its odd derivatives vanish.

\begin{definition}[Taylor expansion and critical diagnostic]
\label{def:qcd-baryon-taylor-critical-diagnostic}
When the thermodynamic continuum pressure is analytic at \(x=0\), write
\begin{equation}
  \frac{p(T,x)}{T^4}
  =
  \sum_{n=0}^{\infty}
  \frac{\chi_n^B(T)}{n!}x^n
  \label{eq:qcd-baryon-pressure-taylor}
\end{equation}
inside its disk of convergence.  The Cauchy--Hadamard radius of this germ is
\begin{equation}
  R(T)
  =
  \left[
  \limsup_{n\to\infty}
  \left|
  \frac{\chi_n^B(T)}{n!}
  \right|^{1/n}
  \right]^{-1}.
  \label{eq:qcd-baryon-cauchy-hadamard-radius}
\end{equation}
\end{definition}

\begin{remark}[What a susceptibility radius establishes]
\label{rem:qcd-baryon-radius-diagnostic-status}
Assume that \(p(T,x)/T^4\) is the germ of a holomorphic function of complex
\(x\) at the origin and has Taylor coefficients
\(\chi_n^B(T)/n!\).  Then \(R(T)\) in
\eqref{eq:qcd-baryon-cauchy-hadamard-radius} is the radius of convergence of
the Taylor series.  The pressure is analytic for \(|x|<R(T)\).  If
independent input shows that the nearest limiting singularity is located at
a positive real point \(x_c\), then \(x_c=R(T)\) and this singularity is the
critical-point candidate seen by the Taylor expansion.  Without that
location input, the nearest singularity may lie at complex \(x\) and the
radius alone does not identify a real critical endpoint.

The first two assertions are the Cauchy--Hadamard theorem applied to the
power series \eqref{eq:qcd-baryon-pressure-taylor}.  A real critical point at
positive \(x_c\) must be a singularity of the same analytic germ after the
thermodynamic and continuum limits have been taken.  If it is independently
known to be the nearest singularity to the origin, its distance from the
origin is the convergence radius.  The final statement is the complementary
case: a power series records distance to the nearest complex obstruction,
whereas the physical critical endpoint is a statement about the real
\((T,\mu_B)\) plane and the limiting Gibbs/KMS states.
\end{remark}

\begin{remark}[Ratio estimators]
\label{rem:qcd-baryon-ratio-estimators}
For a charge-conjugation invariant theory the odd susceptibilities vanish at
\(x=0\), and one often forms even-order estimators
\begin{equation}
  R_{2n}(T)
  =
  \sqrt{
    \frac{(2n+2)(2n+1)\chi_{2n}^B(T)}
         {\chi_{2n+2}^B(T)}
  }.
  \label{eq:qcd-baryon-ratio-estimator}
\end{equation}
If the even Taylor coefficients are eventually positive and have a single
dominant singular scale, these estimators can converge to the radius.  The
positivity and single-scale hypotheses are part of the diagnostic; they are
not consequences of the cumulant definitions alone.
\end{remark}

\section{High-Density Effective Theory and Fermi-Surface Instability}
\label{sec:qcd-high-density-effective-theory}

Real baryon density is difficult nonperturbatively, but asymptotically large
quark chemical potential has a separate weak-coupling control parameter.  The
running coupling is evaluated at a scale of order \(\mu_q\), while the
low-energy excitations live in thin shells around the Fermi surface.  The
resulting effective theory is not an expansion around empty space; it is an
expansion around a filled Fermi sea with soft residual momenta.  The general
finite-regulator Fermi-surface EFT, patch scaling, BCS flow, and
Luttinger-count status discipline are developed in
Section~\ref{sec:fermi-surface-eft-patch-rg}; this section specializes that
framework to weak-coupling dense QCD, where hard-dense-loop gauge response and
color pairing add new dynamical input.

\begin{controlledapproximation}[High-density effective-theory regime]
\label{ca:qcd-hdet-regime}
Let
\[
  \mu_q=\frac{\mu_B}{N_c}.
\]
The high-density effective theory applies to modes whose residual energy and
momentum are bounded by a scale \(\Lambda_{\rm H}\) satisfying
\[
  \Delta,\ T,\ |\ell^\mu|,\ \Lambda_{\rm QCD}
  \ll
  \Lambda_{\rm H}
  \ll
  \mu_q ,
\]
where \(\Delta\) denotes any superconducting gap scale that is being
resolved.  Gauge fields retained in the effective theory have momenta much
smaller than \(\mu_q\).  Modes far from the Fermi surface and antiparticle
branches are integrated out into local operators and into hard-dense-loop
response terms.
\end{controlledapproximation}

\begin{definition}[Patch decomposition near the quark Fermi surface]
\label{def:qcd-hdet-patch-decomposition}
For each unit vector \(\mathbf v\in S^2\), write a quark momentum near the
Fermi surface as
\[
  \mathbf p=\mu_q\mathbf v+\boldsymbol\ell,
  \qquad
  |\boldsymbol\ell|\le \Lambda_{\rm H}.
\]
Decompose the residual momentum into longitudinal and transverse parts,
\[
  \ell_\parallel=\mathbf v\cdot\boldsymbol\ell,
  \qquad
  \boldsymbol\ell_\perp
  =
  \boldsymbol\ell-\mathbf v\,\ell_\parallel.
\]
The patch field \(\psi_{\mathbf v}\) is the positive-energy quark field
whose momenta lie in this shell and whose group velocity is near
\(\mathbf v\).  More invariantly, it is obtained by applying the positive
spectral projector of the massless one-particle Dirac Hamiltonian in the
direction \(\mathbf v\), multiplying by the phase
\(e^{-i\mu_q\mathbf v\cdot\mathbf x}\), and then restricting residual
momenta to the shell.  A finite regulator is obtained by choosing an
angular partition of \(S^2\) whose cells have diameter
\(O(\Lambda_{\rm H}/\mu_q)\); every occurrence of \(\sum_{\mathbf v}\)
below denotes this regulated patch sum, with the continuum angular integral
recovered by refining the partition inside the scale separation
\(\Lambda_{\rm H}\ll\mu_q\).
\end{definition}

\paragraph{Fermi-surface measure and density of states.}
For functions supported in the shell
\(|\boldsymbol\ell|\le\Lambda_{\rm H}\ll\mu_q\),
\[
  \int\frac{d^3p}{(2\pi)^3}f(\mathbf p)
  =
  \frac{\mu_q^2}{(2\pi)^3}
  \int_{S^2}d\Omega_{\mathbf v}
  \int d\ell_\parallel\,
  f(\mu_q\mathbf v+\ell_\parallel\mathbf v)
  +O\!\left(\frac{\Lambda_{\rm H}}{\mu_q}\right)
\]
after angular patches are refined so that transverse residual momenta are
counted once.  The density of states at the Fermi surface, per spin state
and per internal state, is therefore
\[
  N_{\rm one\ spin}(0)=\frac{\mu_q^2}{2\pi^2}.
\]
Including the two spin states of a massless Dirac quark gives
\[
  N_{\rm Dirac}(0)=\frac{\mu_q^2}{\pi^2}.
\]

Write \(\mathbf p=(\mu_q+\ell_\parallel)\mathbf v+\boldsymbol\ell_\perp\)
inside a patch.  To leading order in \(\Lambda_{\rm H}/\mu_q\), the radial
Jacobian is
\[
  p^2dp\,d\Omega_{\mathbf v}
  =
  (\mu_q+\ell_\parallel)^2d\ell_\parallel\,d\Omega_{\mathbf v}
  =
  \mu_q^2
  \left(1+O\!\left(\frac{\Lambda_{\rm H}}{\mu_q}\right)\right)
  d\ell_\parallel\,d\Omega_{\mathbf v}.
\]
The shell decomposition is a partition of momentum space near the Fermi
surface after a choice of angular patches; the patch refinement only avoids
double counting transverse residual momenta.  Integrating the leading
measure over \(S^2\) gives
\[
  \frac{\mu_q^2}{(2\pi)^3}\,4\pi
  =
  \frac{\mu_q^2}{2\pi^2}
\]
per spin and internal state.  A massless Dirac quark has two spin states on
the positive-energy branch, giving the final formula.

\paragraph{Tree-level high-density quark action.}
For a massless quark near the Fermi surface, the positive-energy dispersion
relative to the chemical potential is
\[
  |\mu_q\mathbf v+\boldsymbol\ell|-\mu_q
  =
  \ell_\parallel
  +
  \frac{\boldsymbol\ell_\perp^2}{2\mu_q}
  +O\!\left(\frac{|\boldsymbol\ell|^3}{\mu_q^2}\right).
\]
Consequently the tree-level inverse propagator in a patch is
\[
  G_{\mathbf v}^{-1}(\ell)
  =
  \ell_0-\ell_\parallel
  -\frac{\boldsymbol\ell_\perp^2}{2\mu_q}
  +O\!\left(\frac{|\boldsymbol\ell|^3}{\mu_q^2}\right),
\]
with soft gauge fields introduced by replacing residual derivatives by
covariant derivatives.  The leading local patch Lagrangian has the form
\[
  \mathcal L_{\rm HDET}^{(0)}
  =
  \sum_{\mathbf v}
  \psi_{\mathbf v}^{\dagger}
  \left(
    iD_t+i\mathbf v\cdot\mathbf D
    +\frac{D_\perp^2}{2\mu_q}
    +\cdots
  \right)
  \psi_{\mathbf v},
\]
where \(D_\perp^2=D_iD_i-(\mathbf v\cdot\mathbf D)^2\), and the ellipsis
contains higher powers of residual momentum divided by \(\mu_q\), quark
masses, and local interactions generated by integrating out hard modes.

The expansion follows from
\[
  |\mu_q\mathbf v+\boldsymbol\ell|
  =
  \mu_q
  \sqrt{
    1+\frac{2\ell_\parallel}{\mu_q}
    +\frac{\ell_\parallel^2+\boldsymbol\ell_\perp^2}{\mu_q^2}
  }.
\]
Taylor expansion of the square root gives
\[
  \mu_q+\ell_\parallel+\frac{\boldsymbol\ell_\perp^2}{2\mu_q}
  +O(|\boldsymbol\ell|^3/\mu_q^2).
\]
Subtracting the chemical potential gives the displayed energy relative to
the filled Fermi sea.  The inverse propagator is \(\ell_0\) minus this
relative energy.  In coordinate space, residual momentum acts as
\(-iD\); hence the terms \(-\ell_\parallel\) and
\(-\boldsymbol\ell_\perp^2/(2\mu_q)\) become
\(i\mathbf v\cdot\mathbf D\) and \(D_\perp^2/(2\mu_q)\), with the sign
fixed by the displayed momentum-space inverse propagator.

\paragraph{Hard-dense-loop Debye mass at zero temperature.}
At \(T=0\), for \(N_f\) massless Dirac quarks with common quark chemical
potential \(\mu_q\), the static dense-quark contribution to the electric
screening mass is
\[
  m_{D,\mu}^2
  =
  g^2\sum_{\text{Dirac }f}T_{R_f}\frac{\mu_q^2}{\pi^2}.
\]
For \(SU(N_c)\) fundamentals in the monograph trace convention
\(\operatorname{tr}_{\square}(T^aT^b)=\delta^{ab}\), this becomes
\[
  m_{D,\mu}^2=g^2N_f\frac{\mu_q^2}{\pi^2}.
\]
This is the one-line susceptibility calculation entering the HDL
approximation.  The static electric response is the color susceptibility of
occupied quark states.  At \(T=0\), differentiating the occupation number
\(\Theta(\mu_q-p)\) with respect to a color chemical potential localizes the
response on the Fermi surface:
\[
  -\frac{\partial}{\partial p}\Theta(\mu_q-p)
  =
  \delta(p-\mu_q).
\]
For one spin state in representation \(R\),
\[
  \int\frac{d^3p}{(2\pi)^3}\delta(p-\mu_q)
  =
  \frac{\mu_q^2}{2\pi^2}.
\]
The two positive-energy spin states of a Dirac quark give
\(\mu_q^2/\pi^2\).  Multiplying by the color trace
\(\operatorname{tr}_{R}(T_R^aT_R^b)=T_R\delta^{ab}\), by \(g^2\), and by the
number of flavors gives the displayed result.  Antiquark states are far
from the positive-\(\mu_q\) Fermi surface and do not contribute to this
leading \(T=0\) dense response.

\section{Dense Non-Fermi-Liquid Scaling}
\label{sec:qcd-non-fermi-liquid-self-energy}

The high-density effective theory also controls the quark two-point function
near an unpaired Fermi surface.  A local four-fermion interaction would give a
regular Fermi-liquid expansion in the external energy.  Dense QCD is more
singular because static magnetic gluons are not Debye screened; their
Landau-damped propagator produces a logarithmic derivative of the quark
self-energy.  This is a perturbative high-density statement, not a
nonperturbative theorem about all finite-density QCD states.

\begin{controlledapproximation}[HDL non-Fermi-liquid self-energy regime]
\label{ca:qcd-hdl-non-fermi-liquid-regime}
Work at \(T=0\), in an unpaired high-density patch of
Definition~\ref{def:qcd-hdet-patch-decomposition}, with
\[
  |p_0|,\ |\ell_\parallel|\ll \Lambda_{\rm H}\ll\mu_q,
  \qquad
  g(\mu_q)\ll1 .
\]
Use the trace-delta color convention
\(\operatorname{tr}_{\square}(T^aT^b)=\delta^{ab}\), so that
\begin{equation}
  C_F=T^aT^a=\frac{N_c^2-1}{N_c}
  \label{eq:qcd-trace-delta-fundamental-casimir}
\end{equation}
on a fundamental quark.  Retain the HDL transverse propagator in the
Landau-damped region
\begin{equation}
  |\omega|\ll k_\perp\ll\mu_q,
  \qquad
  D_T(\omega,k_\perp)
  =
  \frac{1}{
    k_\perp^2
    +\frac{\pi}{4}m_{D,\mu}^2\frac{|\omega|}{k_\perp}
  }
  \left(1+o(1)\right),
  \label{eq:qcd-nfl-landau-damped-propagator}
\end{equation}
with \(m_{D,\mu}\) given by the zero-temperature HDL susceptibility
calculation above.  Electric exchange,
non-collinear transverse exchange, and hard matching contributions are kept
only to the extent that they define the finite matching scale
\(\Lambda_{\rm NFL}\); they do not affect the coefficient of the logarithm
derived below.
\end{controlledapproximation}

\begin{definition}[Leading non-Fermi-liquid self-energy]
\label{def:qcd-leading-nfl-self-energy}
Let \(\Sigma(ip_0,\ell_\parallel)\) be the one-particle-irreducible quark
self-energy of a patch field \(\psi_{\mathbf v}\), projected onto the
positive-energy branch and evaluated in Euclidean frequency.  The leading
non-Fermi-liquid part is the nonanalytic term in
\[
  G_{\mathbf v}^{-1}(ip_0,\ell_\parallel)
  =
  ip_0-\ell_\parallel-\Sigma(ip_0,\ell_\parallel)
\]
which is proportional to \(ip_0\log |p_0|\) at
\(\ell_\parallel=0\).  Local analytic terms proportional to \(ip_0\),
\(\ell_\parallel\), or higher powers of residual momentum are scheme
dependent HDET counterterms and are not included in
\(\Sigma_{\rm NFL}\).
\end{definition}

\paragraph{Leading logarithmic cold dense quark self-energy.}
In Controlled Approximation~\ref{ca:qcd-hdl-non-fermi-liquid-regime}, the
on-Fermi-surface Euclidean self-energy has the leading nonanalytic form
\begin{equation}
  \Sigma_{\rm NFL}(ip_0,0)
  =
  ip_0\,\lambda_{\rm NFL}
  \log\frac{\Lambda_{\rm NFL}}{|p_0|}
  +O(ip_0),
  \qquad
  \lambda_{\rm NFL}
  =
  \frac{g^2C_F}{12\pi^2}.
  \label{eq:qcd-leading-nfl-self-energy}
\end{equation}
For \(SU(3)\) in the monograph trace convention,
\begin{equation}
  C_F=\frac83,
  \qquad
  \lambda_{\rm NFL}=\frac{2g^2}{9\pi^2}.
  \label{eq:qcd-leading-nfl-self-energy-su3}
\end{equation}

The one-loop patch self-energy from transverse exchange is obtained by
contracting the patch quark current with the transverse HDL propagator.  The
color projection gives \(C_F\).  Choose the patch normal
\(\mathbf v=\mathbf e_3\).  The leading transverse vertex is
\(gT^a v^i\), and the transverse projector gives
\[
  v^i\left(\delta_{ij}-\frac{k_i k_j}{|\mathbf k|^2}\right)v^j
  =
  1+O\!\left(\frac{k_\parallel^2}{k_\perp^2}\right)
\]
in the Landau-damped region after the internal quark pole restricts
\(|k_\parallel|\lesssim |\omega|\ll k_\perp\).  Thus the singular part is
independent of the detailed spinor wave functions, while analytic terms are
absorbed into local HDET counterterms.

With the external residual momentum set to \(\ell_\parallel=0\), subtract
\(\Sigma(0,0)\) and integrate first over the normal residual momentum of the
internal quark.  For \(p_0>0\), the residue theorem gives the reduced
singular expression
\begin{equation}
  \Sigma_{\rm sing}(ip_0,0)-\Sigma_{\rm sing}(0,0)
  =
  i\,\frac{g^2C_F}{4\pi^2}
  \int_0^{p_0}d\nu\,
  \mathcal L(\nu)
  +O(ip_0),
  \label{eq:qcd-nfl-reduced-patch-integral}
\end{equation}
where \(\nu\) is the Euclidean energy carried by the nearly on-shell
internal quark and \(\mathcal L(\nu)\) is the remaining transverse
Landau-damped logarithm.  The prefactor in
\eqref{eq:qcd-nfl-reduced-patch-integral} is the product of the color
factor, the patch phase-space measure, the transverse-current projection,
and the \(2\pi i\) residue of the normal propagator.  For \(p_0<0\) the same
formula holds with the integral oriented from \(0\) to \(p_0\); this is why
the final answer is proportional to \(ip_0\).

The only nonanalytic part comes from the Landau-damped transverse region.
For fixed \(\nu\), the HDL denominator in
\eqref{eq:qcd-nfl-landau-damped-propagator} cuts off the transverse integral
at
\[
  k_{\perp,{\rm min}}^3\asymp m_{D,\mu}^2|\nu|,
  \qquad
  k_{\perp,{\rm max}}\asymp\mu_q ,
\]
and therefore
\begin{equation}
  \int_{k_{\perp,{\rm min}}}^{k_{\perp,{\rm max}}}
  \frac{dk_\perp}{k_\perp}
  =
  \frac13\log\frac{\Omega_{\rm NFL}}{|\nu|}
  +O(1),
  \qquad
  \Omega_{\rm NFL}\asymp\frac{\mu_q^3}{m_{D,\mu}^2}.
  \label{eq:qcd-nfl-collinear-log}
\end{equation}
The \(O(1)\) part and the precise constant in \(\Omega_{\rm NFL}\) are
absorbed into the matching scale \(\Lambda_{\rm NFL}\).  Equivalently,
\begin{equation}
  \mathcal L(\nu)
  =
  \frac13\log\frac{\Lambda_{\rm NFL}}{|\nu|}
  +O(1).
  \label{eq:qcd-nfl-transverse-log-function}
\end{equation}
Substitution into \eqref{eq:qcd-nfl-reduced-patch-integral} gives
\begin{equation}
  \Sigma_{\rm NFL}(ip_0,0)
  =
  i\,\frac{g^2C_F}{4\pi^2}
  \int_0^{p_0}d\nu\,
  \left(
    \frac13\log\frac{\Lambda_{\rm NFL}}{|\nu|}
  \right)
  +O(ip_0).
  \label{eq:qcd-nfl-reduced-energy-integral}
\end{equation}
The endpoint singularity is integrable and
\[
  \int_0^{p_0}d\nu\,
  \log\frac{\Lambda_{\rm NFL}}{\nu}
  =
  p_0\log\frac{\Lambda_{\rm NFL}}{p_0}
  +p_0
  \qquad (p_0>0).
\]
The second term is analytic and is included in \(O(ip_0)\).  Odd continuation
in \(p_0\), written with \(|p_0|\), gives
\[
  ip_0\,\frac{g^2C_F}{12\pi^2}
  \log\frac{\Lambda_{\rm NFL}}{|p_0|}
  +O(ip_0),
  \]
which is \eqref{eq:qcd-leading-nfl-self-energy}.  The \(SU(3)\)
specialization follows from \eqref{eq:qcd-trace-delta-fundamental-casimir}.

\begin{remark}[Logarithmic vanishing of the patch residue]
\label{rem:qcd-nfl-quark-residue}
Under the assumptions of
the leading logarithmic self-energy calculation above, the coefficient of
\(ip_0\) in the inverse patch propagator behaves as
\begin{equation}
  Z^{-1}(p_0)
  =
  1+\lambda_{\rm NFL}\log\frac{\Lambda_{\rm NFL}}{|p_0|}
  +O(1),
  \label{eq:qcd-nfl-residue}
\end{equation}
and hence the perturbative patch residue satisfies
\begin{equation}
  Z(p_0)\sim
  \left[
    \lambda_{\rm NFL}\log\frac{\Lambda_{\rm NFL}}{|p_0|}
  \right]^{-1}
  \longrightarrow 0
  \qquad
  (p_0\to0)
  \label{eq:qcd-nfl-residue-limit}
\end{equation}
within the unpaired high-density regime.

Insert \eqref{eq:qcd-leading-nfl-self-energy} into the inverse propagator of
Definition~\ref{def:qcd-leading-nfl-self-energy}.  Analytic wave-function
renormalizations change the \(O(1)\) part of
\eqref{eq:qcd-nfl-residue} but cannot remove the logarithm.  Inverting the
coefficient of \(ip_0\) gives \eqref{eq:qcd-nfl-residue-limit}.
The statement is perturbative and patch-local: it identifies the logarithmic
loss of a quasiparticle pole inside the unpaired high-density effective
description, with pairing gaps, confinement-scale physics, and regulator
matching held outside the displayed asymptotic regime.
\end{remark}

\begin{remark}[Relation to the superconducting gap]
\label{rem:qcd-nfl-gap-relation}
The logarithm in \eqref{eq:qcd-leading-nfl-self-energy} is generated by the
same Landau-damped magnetic gluons that produce the leading magnetic pairing
kernel.  It does not change the leading \(1/g\) exponent in
Proposition~\ref{prop:qcd-leading-magnetic-gap-exponent}; it changes
subleading matching data and prefactors.  Thus the phrase
``non-Fermi liquid'' here has a precise meaning: the derivative of the
inverse quark propagator has a logarithmic singularity in the unpaired
high-density effective theory.  It is not a claim that gauge-invariant
observables in every dense phase have been nonperturbatively constructed.
\end{remark}

\paragraph{BCS logarithm from the Fermi surface.}
Consider a specified antisymmetric two-quark channel of the high-density
effective theory and approximate its attractive interaction by a local
four-fermion coupling \(G>0\) inside a shell
\(|\xi|\le\Lambda_{\rm H}\), where
\(\xi=|\mathbf p|-\mu_q\).  Let \(N_{\rm ch}(0)\) be the density of states
in that channel at the Fermi surface, including the relevant spin, color,
and flavor multiplicity.  The zero-temperature mean-field gap equation is
\[
  1
  =
  G N_{\rm ch}(0)
  \int_0^{\Lambda_{\rm H}}
  \frac{d\xi}{\sqrt{\xi^2+\Delta^2}} .
\]
For \(\Delta\ll\Lambda_{\rm H}\),
\[
  \int_0^{\Lambda_{\rm H}}
  \frac{d\xi}{\sqrt{\xi^2+\Delta^2}}
  =
  \log\frac{2\Lambda_{\rm H}}{\Delta}
  +O\!\left(\frac{\Delta^2}{\Lambda_{\rm H}^2}\right),
\]
and hence
\[
  \Delta
  =
  2\Lambda_{\rm H}
  \exp\!\left[-\frac{1}{G N_{\rm ch}(0)}\right]
  \left(1+O\!\left(\frac{\Delta^2}{\Lambda_{\rm H}^2}\right)\right).
\]
In the paired mean-field Hamiltonian, the quasiparticle energy in the shell
is
\[
  E(\xi)=\sqrt{\xi^2+\Delta^2}.
\]
The anomalous expectation value is proportional to
\(\Delta/(2E(\xi))\).  Dividing the self-consistency equation by
\(\Delta\ne0\) gives the displayed integral equation after the angular and
internal sums are collected into \(N_{\rm ch}(0)\).  The integral is exactly
\[
  \int_0^{\Lambda_{\rm H}}\frac{d\xi}{\sqrt{\xi^2+\Delta^2}}
  =
  \operatorname{arsinh}\frac{\Lambda_{\rm H}}{\Delta}
  =
  \log\left(
    \frac{\Lambda_{\rm H}}{\Delta}
    +
    \sqrt{1+\frac{\Lambda_{\rm H}^2}{\Delta^2}}
  \right).
\]
Expanding for \(\Delta/\Lambda_{\rm H}\to0\) gives
\(\log(2\Lambda_{\rm H}/\Delta)+O(\Delta^2/\Lambda_{\rm H}^2)\), and solving
the gap equation gives the stated exponential scale.

\begin{remark}[What is and is not contained in the BCS logarithm]
\label{rem:qcd-bcs-logarithm-status}
The preceding calculation proves the kinematic Fermi-surface
instability for a local attractive channel.  In QCD the microscopic
attraction, the color-flavor structure, and the weak-coupling gap prefactor
also involve electric screening and dynamically screened magnetic gluons.
Those ingredients refine the value of \(\Delta\); they do not remove the
basic logarithmic instability once an attractive channel is present.
\end{remark}

\section{Neutrality Constraints and Gauge Sources}
\label{sec:qcd-neutrality-gauge-sources}

Dense QCD calculations often introduce several quantities called chemical
potentials.  The logically primary distinction is between sources for
gauge-invariant global charges and variables that enforce gauge constraints.
The two have different status already at finite regulator.

\begin{definition}[Finite-regulator global-charge ensemble]
\label{def:qcd-finite-regulator-global-charge-ensemble}
Let \(\mathcal H_{\Lambda}\) be the regulated Hilbert space of a finite
spatial lattice or another finite gauge regulator, let \(H_{\Lambda}\) be
the regulated Hamiltonian, and let \(\mathcal G_{\Lambda}\) be the compact
group of time-independent gauge transformations.  Write
\[
  \Pi_{\rm phys}
  =
  \int_{\mathcal G_{\Lambda}}d\eta\,U(\eta)
\]
for the normalized Haar projector onto gauge-invariant states.  If
\(Q_A\) are mutually commuting conserved global charges which commute with
\(\mathcal G_{\Lambda}\), the finite-regulator grand partition function is
\begin{equation}
  Z_{\Lambda}(\boldsymbol\mu)
  =
  \operatorname{Tr}_{\mathcal H_{\Lambda}}
  \left[
    \Pi_{\rm phys}
    \exp\!\left(
      -\beta H_{\Lambda}
      +\beta\sum_A\mu_A Q_A
    \right)
  \right].
  \label{eq:qcd-global-charge-grand-canonical}
\end{equation}
The intensive charge density conjugate to \(\mu_A\) is
\begin{equation}
  n_A
  =
  \frac{1}{\beta V}
  \frac{\partial}{\partial\mu_A}
  \log Z_{\Lambda}(\boldsymbol\mu),
  \label{eq:qcd-global-charge-density}
\end{equation}
with the thermodynamic and continuum limits taken only after the finite
quantity has been defined.
\end{definition}

\paragraph{Gauss-law annihilation of gauge charge.}
Let \(G_{\Lambda}(f)\) be the infinitesimal generator of a one-parameter
family of gauge transformations \(U(e^{tf})\subset\mathcal G_{\Lambda}\).
Then
\begin{equation}
  G_{\Lambda}(f)\Pi_{\rm phys}=0,
  \qquad
  \Pi_{\rm phys}G_{\Lambda}(f)=0.
  \label{eq:qcd-gauss-projector-annihilation}
\end{equation}
Consequently every finite-regulator state defined by
\eqref{eq:qcd-global-charge-grand-canonical} has zero gauge charge:
\begin{equation}
  \left\langle G_{\Lambda}(f)\right\rangle_{\boldsymbol\mu}=0.
  \label{eq:qcd-gauge-charge-expectation-zero}
\end{equation}
Thus a color charge cannot be assigned an independent physical chemical
potential in the same sense as baryon number or another gauge-invariant
global charge.

The Haar projector satisfies \(U(\eta)\Pi_{\rm phys}=\Pi_{\rm phys}\) and
\(\Pi_{\rm phys}U(\eta)=\Pi_{\rm phys}\) for every
\(\eta\in\mathcal G_{\Lambda}\), by left and right invariance of Haar
measure.  Apply this identity to the one-parameter subgroup \(e^{tf}\) and
differentiate at \(t=0\).  This gives
\eqref{eq:qcd-gauss-projector-annihilation}.  Inserting \(G_{\Lambda}(f)\)
inside the trace defining the expectation value, and using either equality
in \eqref{eq:qcd-gauss-projector-annihilation} together with cyclicity of
the finite trace, gives \eqref{eq:qcd-gauge-charge-expectation-zero}.

\begin{remark}[Meaning of color-neutrality variables]
\label{rem:qcd-color-chemical-potential-language}
In a gauge-fixed high-density calculation one may parametrize a homogeneous
temporal gauge background by Cartan components \(A_0^a\), or equivalently by
symbols often denoted \(\mu_3,\mu_8\).  These symbols are not external
thermodynamic chemical potentials for physical color charge.  They are
coordinates on the gauge-field zero mode left after gauge fixing.  If
\(\Omega_{\rm gf}(A_0)\) is the gauge-fixed finite-volume effective
potential obtained after the nonzero modes have been integrated out, a
homogeneous saddle representing a physical bulk state must obey
\begin{equation}
  \frac{\partial\Omega_{\rm gf}}{\partial A_0^a}=0
  \label{eq:qcd-color-neutrality-stationarity}
\end{equation}
for every retained Cartan component.  Equation
\eqref{eq:qcd-color-neutrality-stationarity} is the saddle form of the
Gauss-law constraint.  Its content is color neutrality; it is not the
definition of a new grand-canonical ensemble.
\end{remark}

\begin{definition}[Neutrality problem for long-range global or gauged charges]
\label{def:qcd-neutrality-problem}
Let \(Q_{\rm ext}\) be a conserved charge which is gauge invariant with
respect to color, for example the electromagnetic charge after coupling QCD
matter to a nondynamical or dynamical \(U(1)\) gauge field.  If the bulk
state is required to be neutral with respect to this charge, introduce a
source \(\mu_{\rm ext}\) in the finite-regulator ensemble and solve
\begin{equation}
  n_{\rm ext}(T,\mu_B,\mu_{\rm ext},\ldots)
  =
  \frac{\partial p}{\partial\mu_{\rm ext}}
  =
  0
  \label{eq:qcd-global-neutrality-equation}
\end{equation}
after all charged species included in the physical problem, such as
electrons in electrically neutral matter, have been included in the
pressure.  If \(Q_{\rm ext}\) is itself gauged and long-ranged in the phase
under consideration, the same condition is the thermodynamic form of the
zero-mode Gauss law: a nonzero uniform charge density would require an
electric field whose energy density is not compatible with an infinite
homogeneous neutral bulk limit.
\end{definition}

\begin{remark}[Ideal CFL is electrically neutral at zero electric source]
\label{rem:qcd-ideal-cfl-neutrality}
Assume \(N_c=N_f=3\), massless \(u,d,s\) quarks, a common quark chemical
potential \(\mu_q\), and the ideal CFL symmetry realization described below:
the gauge-invariant long-distance order preserves the diagonal vector
\(SU(3)\) flavor symmetry.  Let
\begin{equation}
  Q_{\rm em}
  =
  \operatorname{diag}
  \left(
    \frac23,-\frac13,-\frac13
  \right)
  \in\mathfrak{su}(3)_V
  \label{eq:qcd-em-charge-matrix}
\end{equation}
act on the three flavor labels.  Then the electric charge density conjugate
to a source \(\mu_Q Q_{\rm em}\) satisfies
\begin{equation}
  n_Q(T,\mu_B,\mu_Q=0)=0
  \label{eq:qcd-ideal-cfl-electric-neutrality}
\end{equation}
inside the ideal CFL effective theory.  Moreover the CFL orientation
\(U=1\) in
Controlled Approximation~\ref{ca:qcd-cfl-higgs-screening-effective-action}
is invariant under the simultaneous color and flavor transformation generated
by \(Q_{\rm em}\), with the same matrix embedded in a Cartan subalgebra of
color.  The corresponding photon-gluon mixture is the unbroken rotated
\(U(1)\) gauge field of ideal CFL, with the relative sign fixed by the
displayed transformation law \(U\mapsto hUg^{-1}\).

The source \(\mu_Q\) couples to the gauge-invariant vector flavor charge
whose one-particle matrix is \(Q_{\rm em}\).  At \(\mu_Q=0\), the assumed
ideal CFL state is invariant under the diagonal vector \(SU(3)\) flavor
symmetry.  Therefore the expectation value of every traceless generator of
\(\mathfrak{su}(3)_V\) vanishes.  Since
\(\operatorname{tr}Q_{\rm em}=2/3-1/3-1/3=0\), the density conjugate to
\(\mu_Q Q_{\rm em}\) is zero.  This proves
\eqref{eq:qcd-ideal-cfl-electric-neutrality}.

For the second assertion, recall the transformation law of the CFL
orientation coordinate,
\[
  U\longmapsto h\,U\,g^{-1},
\]
where \(h\) is a color gauge transformation and \(g\) is a locked vector
flavor transformation.  With \(U=1\), choosing
\[
  h=e^{i\alpha Q_{\rm em}},
  \qquad
  g=e^{i\alpha Q_{\rm em}}
\]
gives \(hUg^{-1}=1\).  Hence this simultaneous transformation is unbroken.
Equivalently, the infinitesimal action on \(U\) is
\(\delta U=iQ_{\rm color}U-iUQ_{\rm flavor}\), so \(U=1\) is fixed by the
diagonal pair \(Q_{\rm color}=Q_{\rm flavor}=Q_{\rm em}\).
Because the color factor is gauged and the electromagnetic flavor factor is
gauged when QED is included, the massless gauge boson is the corresponding
linear combination of the photon and the two Cartan gluons.  The statement is
independent of the gauge choice used to display \(U=1\); the gauge-invariant
content is the existence of the unbroken charge acting trivially on the CFL
order.
\end{remark}

\begin{remark}[When neutrality becomes dynamical]
\label{rem:qcd-neutrality-beyond-ideal-cfl}
Quark masses, weak equilibrium constraints, finite temperature, boundary
conditions, and possible crystalline or meson-condensed order can break the
ideal assumptions of Remark~\ref{rem:qcd-ideal-cfl-neutrality}.  Then
neutrality is not a slogan attached to a named phase.  It is the coupled
problem consisting of the global neutrality equations
\eqref{eq:qcd-global-neutrality-equation}, the gauge stationarity equations
\eqref{eq:qcd-color-neutrality-stationarity}, and the variational equations
for the order parameters retained in the controlled effective theory.
\end{remark}

\section{Fermi-Surface Stress from Masses and Neutrality}
\label{sec:qcd-fermi-surface-stress}

The ideal CFL limit has equal massless quarks and a common quark chemical
potential.  Real dense matter introduces flavor-dependent stresses through
quark masses, electric neutrality, and weak-equilibrium constraints.  The
first controlled step is kinematic: determine how these deformations shift
the location of the Fermi surface before asking which paired state minimizes
the effective potential.

\begin{definition}[Effective Fermi-surface mismatch]
\label{def:qcd-effective-fermi-surface-mismatch}
Consider two quark species \(i,j\) with masses \(m_i,m_j\) and chemical
potentials \(\mu_i,\mu_j\).  Assume
\[
  m_i^2,\ m_j^2,\ |\mu_i-\mu_j|\,\bar\mu
  \ll
  \bar\mu^2,
  \qquad
  \bar\mu=\frac{\mu_i+\mu_j}{2}.
\]
For momenta in a common patch near \(p=\bar\mu\), define the residual
single-particle energies
\begin{equation}
  \xi_a(p)=\sqrt{p^2+m_a^2}-\mu_a,
  \qquad a=i,j.
  \label{eq:qcd-mass-stressed-xi}
\end{equation}
The average residual energy and half-mismatch are
\begin{equation}
  \bar\xi_{ij}(p)
  =
  \frac{\xi_i(p)+\xi_j(p)}{2},
  \qquad
  \delta_{ij}^{\rm eff}
  =
  \frac{\xi_i(p)-\xi_j(p)}{2}
  \quad\text{evaluated to leading order at }p=\bar\mu .
  \label{eq:qcd-effective-mismatch-definition}
\end{equation}
\end{definition}

\paragraph{Mismatch expansion.}
The mass and chemical-potential contributions to the effective Fermi-surface
mismatch are fixed by the following elementary expansion.  In the regime of
Definition~\ref{def:qcd-effective-fermi-surface-mismatch},
\begin{equation}
  p_{F,a}
  =
  \sqrt{\mu_a^2-m_a^2}
  =
  \mu_a-\frac{m_a^2}{2\mu_a}
  +O\!\left(\frac{m_a^4}{\mu_a^3}\right),
  \label{eq:qcd-mass-shifted-fermi-momentum}
\end{equation}
and
\begin{equation}
  \delta_{ij}^{\rm eff}
  =
  -\frac{\mu_i-\mu_j}{2}
  +
  \frac{m_i^2-m_j^2}{4\bar\mu}
  +O\!\left(
    \frac{m^4}{\bar\mu^3},
    \frac{|\mu_i-\mu_j|\,m^2}{\bar\mu^2}
  \right).
  \label{eq:qcd-effective-mismatch-formula}
\end{equation}
Equivalently, the leading Fermi-momentum splitting is
\begin{equation}
  p_{F,i}-p_{F,j}
  =
  (\mu_i-\mu_j)
  -
  \frac{m_i^2-m_j^2}{2\bar\mu}
  +O\!\left(
    \frac{m^4}{\bar\mu^3},
    \frac{|\mu_i-\mu_j|\,m^2}{\bar\mu^2}
  \right).
  \label{eq:qcd-fermi-momentum-splitting}
\end{equation}
For a strange quark paired with a massless light quark at common chemical
potential \(\mu_q\), the Fermi-momentum splitting scale is
\begin{equation}
  |p_{F,s}-p_{F,\ell}|
  =
  \frac{m_s^2}{2\mu_q}
  +O(m_s^4/\mu_q^3),
  \label{eq:qcd-strange-fermi-surface-stress}
\end{equation}
while the pairwise half-mismatch is
\begin{equation}
  |\delta_{s\ell}^{\rm eff}|
  =
  \frac{m_s^2}{4\mu_q}
  +O(m_s^4/\mu_q^3).
  \label{eq:qcd-strange-half-mismatch}
\end{equation}

The Fermi momentum solves \(\sqrt{p_{F,a}^2+m_a^2}=\mu_a\), giving the first
identity in \eqref{eq:qcd-mass-shifted-fermi-momentum}.  Expanding the square
root gives
\[
  \sqrt{\mu_a^2-m_a^2}
  =
  \mu_a
  \sqrt{1-\frac{m_a^2}{\mu_a^2}}
  =
  \mu_a-\frac{m_a^2}{2\mu_a}
  +O(m_a^4/\mu_a^3).
\]
For the mismatch, expand \eqref{eq:qcd-mass-stressed-xi} at
\(p=\bar\mu\):
\[
  \xi_a(\bar\mu)
  =
  \bar\mu-\mu_a+\frac{m_a^2}{2\bar\mu}
  +O(m_a^4/\bar\mu^3).
\]
Taking half the difference between \(a=i\) and \(a=j\) gives
\eqref{eq:qcd-effective-mismatch-formula}; keeping the dependence of the
denominator on \(\mu_a\) rather than \(\bar\mu\) produces the displayed
mixed remainder.  Subtracting
\eqref{eq:qcd-mass-shifted-fermi-momentum} for \(i\) and \(j\) gives
\eqref{eq:qcd-fermi-momentum-splitting}.  The strange-light specializations
set \(m_\ell=0\) and \(\mu_s=\mu_\ell=\mu_q\).

\begin{controlledapproximation}[Two-species stressed BCS comparison]
\label{ca:qcd-two-species-stressed-bcs}
To isolate the local pairing stress, consider two species with a common
density of states \(N(0)\), a momentum-independent attractive interaction in
a shell, zero temperature, and half-mismatch \(\delta\).  Let \(\Delta_0\) be
the BCS gap at \(\delta=0\), computed in the same cutoff scheme.  Compare only
the homogeneous paired state with equal paired Fermi momenta to the unpaired
normal state with split Fermi surfaces.  This is a local diagnostic for the
onset of stress; it is not a complete variational analysis of CFL, gapless
CFL, meson-condensed phases, or crystalline phases.
\end{controlledapproximation}

\paragraph{Clogston scale in the controlled two-species model.}
In Controlled Approximation~\ref{ca:qcd-two-species-stressed-bcs}, the
zero-temperature grand-potential difference between the homogeneous paired
state and the split normal state has the leading form
\begin{equation}
  \Omega_{\rm pair}-\Omega_{\rm normal}
  =
  -\frac12N(0)\Delta_0^2
  +
  N(0)\delta^2
  +O(\Delta_0^4/\Lambda_{\rm H}^2).
  \label{eq:qcd-clogston-potential-difference}
\end{equation}
Thus the homogeneous paired state ceases to win against the split normal
state in this model when
\begin{equation}
  |\delta|>\delta_{\rm C},
  \qquad
  \delta_{\rm C}=\frac{\Delta_0}{\sqrt2}.
  \label{eq:qcd-clogston-scale}
\end{equation}

At \(\delta=0\), the BCS condensation energy in the constant-density shell is
\[
  \Omega_{\rm pair}(\delta=0)-\Omega_{\rm normal}(\delta=0)
  =
  -\frac12N(0)\Delta_0^2
  +O(\Delta_0^4/\Lambda_{\rm H}^2),
\]
which follows by substituting the gap equation into the mean-field energy
functional: the logarithmic part of the interaction energy cancels the
logarithmic part of the quasiparticle energy, leaving the displayed finite
term.  A normal state with half-mismatch \(\delta\) fills one species up to
energy \(+\delta\) relative to the average Fermi surface and empties the
other down to \(-\delta\).  With constant density of states, the energy
lowering of the split normal state relative to the unsplit normal state is
\[
  -2N(0)\int_0^{|\delta|}\xi\,d\xi
  =
  -N(0)\delta^2,
\]
where the factor \(2\) counts the filled particle interval for one species
and the emptied hole interval for the other.  Therefore, when the paired
state remains unsplit, its grand potential relative to the split normal state
is shifted upward by \(N(0)\delta^2\), giving
\eqref{eq:qcd-clogston-potential-difference}.  Setting the leading
difference to zero gives \eqref{eq:qcd-clogston-scale}.

\begin{remark}[Use of the stress scale in dense QCD]
\label{rem:qcd-stress-scale-status}
Equations \eqref{eq:qcd-strange-fermi-surface-stress} and
\eqref{eq:qcd-clogston-scale} should be used with their hypotheses attached.
The scale \(m_s^2/(2\mu_q)\) is the leading strange-light Fermi-momentum
splitting, while \(m_s^2/(4\mu_q)\) is the pairwise half-mismatch in the
two-species reduction.  Full dense QCD also requires the neutrality equations
of Section~\ref{sec:qcd-neutrality-gauge-sources}, the coupled gap equations
in all pairing channels, and the possibility that the minimizing state is
not spatially homogeneous.
\end{remark}

\section{Color Superconductivity and CFL Order}
\label{sec:qcd-color-superconductivity-cfl}

At asymptotically large \(\mu_B\), quarks near the Fermi surface interact
through channels that can be attractive.  The BCS mechanism is then a
controlled weak-coupling instability of the Fermi surface, though the
precise gap prefactor is sensitive to magnetic gluon exchange and screening.

\paragraph{Color factor of one-gluon exchange between two quarks.}
Let the \(SU(N_c)\) fundamental generators be normalized as throughout the
monograph,
\[
  \operatorname{tr}_{\square}(T^aT^b)=\delta^{ab}.
\]
In the tensor product
\[
  \square\otimes\square
  =
  \operatorname{Sym}^2\square
  \oplus
  \wedge^2\square ,
\]
the color operator \(T_1^aT_2^a\) has eigenvalues
\[
  \left(T_1\cdot T_2\right)_{\operatorname{Sym}^2\square}
  =
  \frac{N_c-1}{N_c},
  \qquad
  \left(T_1\cdot T_2\right)_{\wedge^2\square}
  =
  -\frac{N_c+1}{N_c}.
\]
With the Born-kernel sign convention in which the static exchange energy is
proportional to
\[
  \mathcal V_R(\mathbf q)
  =
  \frac{g^2}{\mathbf q^2}
  \left(T_1\cdot T_2\right)_R ,
\]
the two-index antisymmetric channel is attractive and the symmetric channel
is repulsive.  For \(SU(3)\), the attractive channel is
\(\wedge^2\square=\overline{\square}\), with color factor \(-4/3\), while
the symmetric sextet has color factor \(2/3\).

Let \(C_2(R)\) denote the quadratic Casimir in the present trace convention.
On an irreducible summand \(R\subset\square\otimes\square\),
\[
  (T_1^a+T_2^a)(T_1^a+T_2^a)=C_2(R)\mathbf 1 .
\]
Since each tensor factor is fundamental,
\[
  2T_1^aT_2^a=C_2(R)-2C_2(\square).
\]
The trace-delta convention doubles the more common half-trace Casimirs, so
\[
  C_2(\square)=\frac{N_c^2-1}{N_c},
\]
\[
  C_2(\operatorname{Sym}^2\square)
  =
  \frac{2(N_c+2)(N_c-1)}{N_c},
  \qquad
  C_2(\wedge^2\square)
  =
  \frac{2(N_c-2)(N_c+1)}{N_c}.
\]
Substitution gives
\[
  \frac12\left[
    \frac{2(N_c+2)(N_c-1)}{N_c}
    -\frac{2(N_c^2-1)}{N_c}
  \right]
  =
  \frac{N_c-1}{N_c},
\]
and
\[
  \frac12\left[
    \frac{2(N_c-2)(N_c+1)}{N_c}
    -\frac{2(N_c^2-1)}{N_c}
  \right]
  =
  -\frac{N_c+1}{N_c}.
\]
The sign statement follows from the displayed exchange-energy convention.

\begin{controlledapproximation}[Leading-log magnetic gap problem]
\label{ca:qcd-leading-log-magnetic-gap}
Work at \(T=0\), \(\mu_q\gg\Lambda_{\rm QCD}\), and
\(g(\mu_q)\ll1\) in the high-density effective theory.  Retain the
attractive two-index antisymmetric quark-quark channel and the HDL-resummed
soft gluon propagator.  The leading logarithm in the gap equation is the
part generated by nearly collinear transverse magnetic exchange with
frequency transfer
\[
  \omega=p_0-q_0
\]
and transverse momentum \(k_\perp\) satisfying
\[
  |\omega|\ll k_\perp\ll \mu_q .
\]
In this region the Euclidean transverse propagator has the form
\begin{equation}
  D_T(\omega,k_\perp)
  =
  \frac{1}{
    k_\perp^2
    +\frac{\pi}{4}m_{D,\mu}^2\frac{|\omega|}{k_\perp}
  }
  \left(1+o(1)\right),
  \label{eq:qcd-hdl-transverse-landau-damped}
\end{equation}
where \(m_{D,\mu}\) is the dense HDL Debye mass computed above.  Electric exchange
and non-collinear magnetic exchange are not discarded; in the leading-log
problem they contribute to the matching scale and to the prefactor, not to
the coefficient of the \(1/g\) exponent.
\end{controlledapproximation}

\paragraph{HDL magnetic leading-log gap equation.}
In the regime specified by
Controlled Approximation~\ref{ca:qcd-leading-log-magnetic-gap}, the gap in the
attractive antisymmetric quark-quark channel obeys, to leading logarithmic
accuracy,
\begin{equation}
  \Delta(p_0)
  =
  \lambda_{\rm mag}
  \int_0^{\Lambda_{\rm M}}
  dq_0\,
  \frac{\Delta(q_0)}{\sqrt{q_0^2+\Delta(q_0)^2}}
  \log\frac{\Lambda_{\rm M}}{|p_0-q_0|}
  +\mathcal R(p_0),
  \label{eq:qcd-hdl-leading-log-gap-equation}
\end{equation}
where \(\mathcal R\) has no additional large collinear logarithm.  The
leading-log coefficient is
\begin{equation}
  \lambda_{\rm mag}
  =
  \frac{g^2 C_{\rm qq}}{12\pi^2},
  \qquad
  C_{\rm qq}
  =
  -\left(T_1\cdot T_2\right)_{\wedge^2\square}
  =
  \frac{N_c+1}{N_c}.
  \label{eq:qcd-magnetic-gap-lambda-general}
\end{equation}
For \(SU(3)\) in the monograph trace convention,
\begin{equation}
  C_{\rm qq}=\frac43,
  \qquad
  \lambda_{\rm mag}=\frac{g^2}{9\pi^2}.
  \label{eq:qcd-magnetic-gap-lambda-su3}
\end{equation}
The scale \(\Lambda_{\rm M}\) is a matching scale of order \(\mu_q\) times
powers of \(g\) and numerical constants fixed only by next-to-leading
matching.  It is not determined by the leading-log equation alone.

The gap equation is the off-diagonal Schwinger--Dyson equation in
Nambu--Gorkov variables, restricted to the shell of
Definition~\ref{def:qcd-hdet-patch-decomposition}.  The radial residual
momentum integral across the Fermi surface gives the BCS denominator
\[
  \int d\ell_\parallel\,
  \frac{1}{q_0^2+\ell_\parallel^2+\Delta(q_0)^2}
  =
  \frac{\pi}{\sqrt{q_0^2+\Delta(q_0)^2}},
\]
up to terms without a large logarithm.  The angular integral is dominated
by small scattering angle, equivalently by transverse momentum transfer
\[
  k_\perp\simeq \mu_q\theta .
\]
For transverse gluons the HDL denominator
\eqref{eq:qcd-hdl-transverse-landau-damped} cuts off the collinear
singularity in the infrared at
\[
  k_{\perp,{\rm min}}^3
  \asymp
  m_{D,\mu}^2|\omega|,
\]
while the upper end is fixed by the edge of the patch or by hard matching,
\(k_{\perp,{\rm max}}\asymp\mu_q\).  Thus the logarithmic part of the
transverse integral is
\[
  \int_{k_{\perp,{\rm min}}}^{k_{\perp,{\rm max}}}
  \frac{dk_\perp}{k_\perp}
  =
  \frac13
  \log\frac{\Omega_{\rm M}}{|\omega|}
  +O(1),
  \qquad
  \Omega_{\rm M}\asymp \frac{\mu_q^3}{m_{D,\mu}^2},
\]
where the \(O(1)\) and the precise constant multiplying
\(\Omega_{\rm M}\) are not leading-log data.  The factor \(1/3\) is kept
explicitly in the coefficient below.  After this factor has been absorbed
into \(\lambda_{\rm mag}\), the remaining energy matching scale is denoted
\(\Lambda_{\rm M}\) in \eqref{eq:qcd-hdl-leading-log-gap-equation}.

The remaining coefficient is fixed by the Fermi-surface measure, the
spin-current projection for transverse exchange, and the color projection.
The shell measure and the radial propagator integral give the common
prefactor \(g^2C_{\rm qq}/(4\pi^2)\) multiplying the transverse collinear
integral.  The Landau-damped magnetic integral contributes only one third
of the logarithm of the energy transfer,
\[
  \int\frac{dk_\perp}{k_\perp}
  =
  \frac13\log\frac{1}{|\omega|}
  +\text{terms independent of }|\omega|
\]
at leading logarithmic accuracy.  Thus the coefficient of the logarithmic
energy integral is
\[
  \frac{g^2C_{\rm qq}}{4\pi^2}\cdot\frac13
  =
  \frac{g^2 C_{\rm qq}}{12\pi^2}.
\]
The color factor is not an independent convention: by the two-quark color
calculation above,
\[
  C_{\rm qq}
  =
  -\left(T_1\cdot T_2\right)_{\wedge^2\square}
  =
  \frac{N_c+1}{N_c}.
\]
For \(N_c=3\) this gives \(C_{\rm qq}=4/3\), hence
\(\lambda_{\rm mag}=g^2/(9\pi^2)\).  Terms from screened electric exchange,
from non-collinear transverse exchange, and from quark self-energy do not
change this leading collinear logarithm; they are included in
\(\mathcal R\) and in the matching scale.

\begin{proposition}[Leading exponent of the magnetic color-superconducting gap]
\label{prop:qcd-leading-magnetic-gap-exponent}
Consider the leading-log integral equation
\begin{equation}
  \Delta(p)
  =
  \lambda
  \int_{\Delta_0}^{\Lambda}
  \frac{dq}{q}\,
  \Delta(q)\,
  \log\frac{\Lambda}{\max(p,q)},
  \qquad
  \Delta_0\ll p,q\ll\Lambda,
  \label{eq:qcd-abstract-leading-log-gap-equation}
\end{equation}
with \(\lambda>0\).  If the solution is nontrivial and varies only on the
logarithmic scale, then
\begin{equation}
  \log\frac{\Lambda}{\Delta_0}
  =
  \frac{\pi}{2\sqrt{\lambda}}
  +O(1).
  \label{eq:qcd-leading-log-gap-general-exponent}
\end{equation}
Applying this to the \(SU(3)\) HDL magnetic kernel in the monograph
trace-delta convention gives the leading exponent
\begin{equation}
  \Delta_0
  =
  \Lambda_{\rm M}
  \exp\!\left[-\frac{3\pi^2}{2g}+O(1)\right].
  \label{eq:qcd-leading-magnetic-gap-su3-trace-delta}
\end{equation}
Equivalently, if \(g_{\rm ht}=\sqrt2\,g\) denotes the coupling in the
common half-trace convention, the exponent is
\[
  \exp\!\left[-\frac{3\pi^2}{\sqrt2\,g_{\rm ht}}\right].
\]
\end{proposition}

\begin{proof}
Set
\[
  x=\log\frac{\Lambda}{p},
  \qquad
  y=\log\frac{\Lambda}{q},
  \qquad
  X=\log\frac{\Lambda}{\Delta_0}.
\]
The leading-log kernel becomes
\[
  \log\frac{\Lambda}{\max(p,q)}=\min(x,y).
\]
Equation~\eqref{eq:qcd-abstract-leading-log-gap-equation} becomes
\[
  \Delta(x)
  =
  \lambda
  \int_0^X dy\,\min(x,y)\,\Delta(y)
  =
  \lambda
  \left[
    \int_0^x dy\,y\,\Delta(y)
    +
    x\int_x^Xdy\,\Delta(y)
  \right].
\]
Differentiating once gives
\[
  \Delta'(x)=\lambda\int_x^Xdy\,\Delta(y),
\]
and differentiating again gives the ordinary differential equation
\[
  \Delta''(x)+\lambda\Delta(x)=0.
\]
The integral equation imposes
\[
  \Delta(0)=0,
  \qquad
  \Delta'(X)=0.
\]
The lowest nontrivial solution is therefore
\[
  \Delta(x)=A\sin(\sqrt{\lambda}\,x),
  \qquad
  \cos(\sqrt{\lambda}\,X)=0,
\]
so
\[
  X=\frac{\pi}{2\sqrt{\lambda}}
\]
at leading logarithmic accuracy.  Higher nodes give smaller gaps and do not
describe the leading instability.  Substituting
\(\lambda=\lambda_{\rm mag}=g^2/(9\pi^2)\) from
\eqref{eq:qcd-magnetic-gap-lambda-su3} gives
\[
  \frac{\pi}{2\sqrt{\lambda_{\rm mag}}}
  =
  \frac{3\pi^2}{2g}.
\]
The \(O(1)\) in \eqref{eq:qcd-leading-magnetic-gap-su3-trace-delta}
records terms outside leading-log accuracy.  The power of \(g\) multiplying
\(\mu_q\) in the prefactor and the numerical constant multiplying it are
next-to-leading data, not consequences of this proposition.
\end{proof}

\paragraph{Exchange symmetry and the CFL tensor.}
For a local \(s\)-wave spin-zero pairing of two left-handed Weyl quarks,
the coefficient multiplying
\[
  q^{a i\alpha}_L q^{b j\beta}_L\epsilon_{\alpha\beta}
\]
is symmetric under the simultaneous interchange
\((a,i)\leftrightarrow(b,j)\).  Therefore a pair in the attractive
color-antisymmetric channel must also be antisymmetric in flavor.  For
\(N_c=N_f=3\), the nondegenerate antisymmetric tensor
\[
  \epsilon^{abA}\epsilon_{ijA}
\]
is invariant precisely under the diagonal subgroup obtained by identifying
color and flavor \(SU(3)\) rotations.

The fermion fields anticommute, while
\(\epsilon_{\alpha\beta}\) is antisymmetric.  Hence
\[
  q^{b j\alpha}_L q^{a i\beta}_L\epsilon_{\alpha\beta}
  =
  -q^{a i\beta}_L q^{b j\alpha}_L\epsilon_{\alpha\beta}
  =
  q^{a i\alpha}_L q^{b j\beta}_L\epsilon_{\alpha\beta},
\]
where the last equality relabels \(\alpha\leftrightarrow\beta\) and uses
\(\epsilon_{\beta\alpha}=-\epsilon_{\alpha\beta}\).  Thus the color-flavor
coefficient must be symmetric under simultaneous exchange.  Since the
two-quark color calculation above identifies the attractive color channel as
antisymmetric, the flavor tensor must be antisymmetric as well.

Let \(U\in SU(3)_{\rm color}\) and \(V\in SU(3)_{\rm flavor}\).  The tensor
\(\epsilon^{abA}\epsilon_{ijA}\) may be viewed as the identity map from the
color \(\overline{\square}\) carried by \(\wedge^2\square_{\rm color}\) to
the flavor \(\overline{\square}\) carried by \(\wedge^2\square_{\rm flavor}\).
It is preserved exactly when these two \(\overline{\square}\) actions agree,
which is the diagonal \(SU(3)\) condition after the chosen color-flavor
identification.  This is the color-flavor locking statement.

\begin{definition}[CFL condensate data]
\label{def:qcd-cfl-condensate-data}
For \(N_c=N_f=3\), color-flavor-locked Higgs data are a source-selected or
gauge-fixed representative of the gauge-covariant pairing tensor
\[
  \Phi^{ab}_{L,ij}
  =
  \langle q^a_{L i} C q^b_{L j}\rangle_J
  =
  \Delta_L\,\epsilon^{abA}\epsilon_{ijA},
  \qquad
  \Phi^{ab}_{R,ij}
  =
  \langle q^a_{R i} C q^b_{R j}\rangle_J
  =
  \Delta_R\,\epsilon^{abA}\epsilon_{ijA},
\]
with color indices \(a,b\), flavor indices \(i,j\), and spinor charge
conjugation matrix \(C\).  The subscript \(J\) means that an infinitesimal
pairing source and the infinite-volume limit are used before the source is
removed, exactly as in any spontaneous-order definition.  The tensor is not a
gauge-invariant local operator.  Its gauge-invariant content is the Higgs
pattern of the color gauge field, the residual global diagonal symmetry, and
the long-distance order of color-singlet composites built from the pairing
tensor.  The displayed tensor locks color and flavor: simultaneous color and
flavor rotations preserving the same diagonal \(SU(3)\) leave the
representative invariant.
\end{definition}

\begin{definition}[Gauge-invariant CFL diagnostics]
\label{def:qcd-cfl-gauge-invariant-diagnostics}
Let
\[
  \varphi_L\in
  \overline{\square}_{\rm color}\otimes
  \overline{\square}_{L},
  \qquad
  \varphi_R\in
  \overline{\square}_{\rm color}\otimes
  \overline{\square}_{R}
\]
denote the antisymmetric spin-zero diquark composite obtained from two
left-handed or two right-handed quark fields:
\begin{equation}
  (\varphi_L)^A{}_{I}(x)
  =
  \frac12
  \epsilon^{A}{}_{ab}\epsilon_{I}{}^{ij}\,
  q^a_{L i}(x)Cq^b_{L j}(x),
  \qquad
  (\varphi_R)^A{}_{\bar I}(x)
  =
  \frac12
  \epsilon^{A}{}_{ab}\epsilon_{\bar I}{}^{ij}\,
  q^a_{R i}(x)Cq^b_{R j}(x).
  \label{eq:qcd-cfl-diquark-varphi}
\end{equation}
Here \(A\) is an anti-fundamental color index, \(I\) is an
anti-fundamental \(SU(3)_L\) flavor index, and \(\bar I\) is an
anti-fundamental \(SU(3)_R\) flavor index.  Each quark has baryon charge
\(1/3\), hence \(\varphi_L\) and \(\varphi_R\) have baryon charge \(2/3\).

The following operators are gauge invariant and carry the physical symmetry
information of the CFL phase:
\begin{equation}
  \mathcal B_L
  =
  \frac1{3!}
  \epsilon_{ABC}\epsilon^{IJK}
  (\varphi_L)^A{}_{I}(\varphi_L)^B{}_{J}(\varphi_L)^C{}_{K},
  \qquad
  \mathcal B_R
  =
  \frac1{3!}
  \epsilon_{ABC}\epsilon^{\bar I\bar J\bar K}
  (\varphi_R)^A{}_{\bar I}(\varphi_R)^B{}_{\bar J}
  (\varphi_R)^C{}_{\bar K},
  \label{eq:qcd-cfl-baryon-superfluid-composites}
\end{equation}
and
\begin{equation}
  \Sigma^{I}{}_{\bar J}
  =
  (\varphi_L^\dagger)^{I}{}_{A}
  (\varphi_R)^A{}_{\bar J}.
  \label{eq:qcd-cfl-chiral-composite}
\end{equation}
The operators \(\mathcal B_L,\mathcal B_R\) are color and flavor singlets
with baryon charge \(2\).  The operator \(\Sigma\) is color invariant,
baryon neutral, and transforms as
\[
  \Sigma\longmapsto g_L\,\Sigma\,g_R^{-1}
  \qquad
  (g_L,g_R)\in SU(3)_L\times SU(3)_R .
\]
Finally, for a path \(\gamma\) from \(y\) to \(x\), the Wilson-line
completion
\begin{equation}
  G_L^{I}{}_{J}(x,y;\gamma)
  =
  \left\langle
  (\varphi_L^\dagger)^{I}{}_{A}(x)\,
  \left[W_{\overline{\square}}(\gamma)\right]^{A}{}_{B}\,
  (\varphi_L)^B{}_{J}(y)
  \right\rangle
  \label{eq:qcd-cfl-wilson-line-diquark-correlator}
\end{equation}
is gauge invariant but path dependent.  It is not a new local order
parameter; it is a gauge-invariant way to test the long-distance order
encoded by a gauge-covariant diquark field.
\end{definition}

\begin{proposition}[Gauge averaging and physical CFL order]
\label{prop:qcd-cfl-gauge-invariant-order}
In a finite-volume regularized gauge theory with a gauge-invariant action,
measure, boundary condition, and source, the expectation value of any local
operator transforming in a nontrivial irreducible representation of the
color gauge group is zero.  Consequently the gauge-covariant diquark
one-point function \(\langle\varphi_L\rangle\) is not itself a physical
local order parameter.  The CFL symmetry realization is instead detected by
the gauge-invariant composites of
Definition~\ref{def:qcd-cfl-gauge-invariant-diagnostics}: long-distance
order in \(\mathcal B_{L,R}\) diagnoses baryon superfluid order, while
long-distance order in \(\Sigma\) diagnoses
\[
  SU(3)_L\times SU(3)_R\longrightarrow SU(3)_{\rm diagonal}.
\]
\end{proposition}

\begin{proof}
Let \(O\) be a local operator whose color indices span a finite-dimensional
representation \(\rho\) of \(SU(3)_{\rm color}\).  Gauge invariance of the
regularized functional integral gives, for every constant color
transformation \(h\in SU(3)\),
\[
  \langle O\rangle=\rho(h)\langle O\rangle .
\]
Averaging over \(h\) with normalized Haar measure gives
\[
  \langle O\rangle
  =
  \int_{SU(3)}dh\,\rho(h)\langle O\rangle .
\]
The Haar average is the projection onto the invariant subspace of
\(\rho\).  If \(\rho\) has no singlet component, this projection is zero.
The diquark field \(\varphi_L\) carries the
\(\overline{\square}_{\rm color}\) representation, so its gauge-invariant
finite-volume expectation vanishes unless one first fixes a gauge or uses a
gauge-covariant source-selected representative as in
Definition~\ref{def:qcd-cfl-condensate-data}.

The composites in
Definition~\ref{def:qcd-cfl-gauge-invariant-diagnostics} are different.
For \(\mathcal B_L\), the color indices are contracted with
\(\epsilon_{ABC}\), and \(\det h=1\) for \(h\in SU(3)\), so the color
factor is invariant.  The flavor contraction is likewise invariant under
\(SU(3)_L\), and the baryon charge is
\[
  3\cdot\frac23=2 .
\]
For \(\mathcal B_R\), the right-handed color indices are contracted by the
same \(\epsilon_{ABC}\) tensor and the flavor indices by the right-handed
\(\epsilon^{ijk}\), so the operator is again an \(SU(3)_{\rm color}\)
singlet of baryon charge \(2\).  Therefore a nonzero source-selected
long-distance limit of the two-point function of \(\mathcal B_L\) or
\(\mathcal B_R\) is a gauge-invariant statement that \(U(1)_B\) is
spontaneously broken.

For \(\Sigma\), the color anti-fundamental index of \(\varphi_R\) is
contracted with the conjugate color index of \(\varphi_L^\dagger\), so the
operator is color invariant.  Its baryon charge is
\[
  -\frac23+\frac23=0.
\]
Since \(\varphi_L^\dagger\) transforms in the fundamental of \(SU(3)_L\)
and \(\varphi_R\) in the anti-fundamental of \(SU(3)_R\), \(\Sigma\)
transforms as \(g_L\Sigma g_R^{-1}\).  A source-selected expectation
proportional to the identity leaves only \(g_L=g_R\), which is the diagonal
chiral subgroup.  The Wilson-line correlator
\eqref{eq:qcd-cfl-wilson-line-diquark-correlator} is gauge invariant
because the endpoint transformations of the Wilson line cancel the color
transformations of the two diquark insertions.  Its path dependence is the
price of using a nonlocal completion of a gauge-covariant field, not an
additional physical symmetry.
\end{proof}

\begin{remark}[Physical CFL Goldstone count]
\label{rem:qcd-cfl-goldstone-count}
Assume \(N_c=N_f=3\), ignore the anomalous axial \(U(1)\), and suppose the
infinite-volume dense state has the gauge-invariant CFL long-distance order
described in
Definition~\ref{def:qcd-cfl-gauge-invariant-diagnostics} and
Proposition~\ref{prop:qcd-cfl-gauge-invariant-order}.  The physical global
symmetry breaking pattern is
\[
  SU(3)_L\times SU(3)_R\times U(1)_B
  \longrightarrow
  SU(3)_{\rm diagonal},
\]
and the number of physical Nambu--Goldstone modes before quark-mass effects
are included is
\[
  (8+8+1)-8=9.
\]

The color \(SU(3)\) is gauged.  Directions that rotate the condensate by a
pure color transformation are Higgs directions of the gauge field, not
generators of physical global particles.  The remaining exact global
generators, in the massless and anomaly-idealized limit stated in the
proposition, span
\[
  \mathfrak{su}(3)_L\oplus\mathfrak{su}(3)_R\oplus\mathfrak u(1)_B .
\]
The CFL tensor identifies a diagonal \(\mathfrak{su}(3)\) subgroup with the
unbroken simultaneous flavor rotation, as shown in the exchange-symmetry
calculation above.  The dimension of the broken global coset is therefore
\(8+8+1-8=9\).  The conclusion is only a group-theoretic consequence of the
source-selected CFL order.  Establishing which region of QCD realizes this
order requires control of dense QCD, where direct lattice importance sampling
is obstructed by the finite-density sign-problem mechanism described above.
\end{remark}

\begin{definition}[Gauge-invariant static screening diagnostics in dense QCD]
\label{def:qcd-gauge-invariant-static-screening}
Let \(x=(0,\mathbf x)\), \(y=(0,\mathbf y)\), and let \(\gamma\) be a
spatial path from \(y\) to \(x\).  For the canonically normalized color
gauge potential \(a_\mu=a_\mu^aT^a\), with field strength
\(\mathfrak f_{\mu\nu}=\mathfrak f_{\mu\nu}^aT^a\), define the Wilson-line
completed static electric and magnetic correlators
\[
  C_E(r;\gamma)
  =
  \sum_{i=1}^3
  \left\langle
  \operatorname{tr}\!\left[
    \mathfrak f_{0i}(x)\,
    W_{\square}(\gamma)\,
    \mathfrak f_{0i}(y)\,
    W_{\square}(\gamma)^{-1}
  \right]\right\rangle_{\rm conn},
\]
\begin{equation}
  C_M(r;\gamma)
  =
  \frac12\sum_{i,j=1}^3
  \left\langle
  \operatorname{tr}\!\left[
    \mathfrak f_{ij}(x)\,
    W_{\square}(\gamma)\,
    \mathfrak f_{ij}(y)\,
    W_{\square}(\gamma)^{-1}
  \right]\right\rangle_{\rm conn},
  \qquad
  r=|\mathbf x-\mathbf y|.
  \label{eq:qcd-cfl-screening-correlators}
\end{equation}
The displayed correlators are bare nonlocal composites.  A continuum
screening observable must first specify an operator scheme
\(\mathcal S\): the operator sector, the adjoint line representation, the path
family \(\{\gamma_r\}\), the endpoint and field-strength renormalizations,
and the subtraction or ratio used to remove the Wilson-line self-energy.  One
convenient notation is
\begin{equation}
  C_X^{\mathcal S}(r)
  =
  Z_{X,\rm end}^{-2}Z_{X,\mathfrak f}^{-2}
  \exp\!\left[\delta m_{\rm line}\ell(\gamma_r)\right]\,
  C_X^{\rm bare}(r;\gamma_r),
  \qquad X=E,M,
  \label{eq:qcd-cfl-renormalized-screening-correlator}
\end{equation}
with all factors evaluated in the chosen regulator and renormalization
scheme.  Equivalently, one may divide by a reference Wilson-line observable
with the same long segment.  What matters is that the line self-energy and
endpoint factors are not left inside the extracted distance exponent.  The
static electric and magnetic screening exponents are then the leading
large-\(r\) exponential decay rates of the renormalized observables, when they
exist after the thermodynamic and continuum limits,
\begin{equation}
  M_E^{\mathcal S}=-\lim_{r\to\infty}\frac1r\log |C_E^{\mathcal S}(r)|,
  \qquad
  M_M^{\mathcal S}=-\lim_{r\to\infty}\frac1r\log |C_M^{\mathcal S}(r)|,
  \label{eq:qcd-cfl-screening-exponents}
\end{equation}
for the chosen scheme \(\mathcal S\).  Their transfer-matrix interpretation is
the lowest state in the sector selected by the nonlocal operator: field
strength insertions, adjoint static line, endpoint operators, and path family.
That state can be a gluelump-like, static-source, multiparticle, or other
composite screening state.  In weakly coupled phases a gauge-fixed
long-distance effective action may be matched to these exponents, but the
matching requires dynamical input beyond symmetry.
\end{definition}

\begin{controlledapproximation}[CFL Higgs screening effective action]
\label{ca:qcd-cfl-higgs-screening-effective-action}
In a dense state with the CFL long-distance order of
Definition~\ref{def:qcd-cfl-gauge-invariant-diagnostics}, introduce a
local gauge-covariant coordinate \(U(x)\in SU(3)\) for the color-flavor
locked orientation.  Under a color gauge transformation \(h(x)\) and a
global locked flavor transformation \(g\), it transforms as
\[
  U(x)\longmapsto h(x)U(x)g^{-1}.
\]
The leading static screening part of the long-distance effective action is
\begin{equation}
  \mathcal L_{\rm H}
  =
  \frac{F_H^2}{2}
  \operatorname{tr}\!\left[(D_0U)^\dagger D_0U\right]
  -
  \frac{F_H^2v_H^2}{2}
  \operatorname{tr}\!\left[(D_iU)^\dagger D_iU\right],
  \label{eq:qcd-cfl-higgs-screening-action}
\end{equation}
where
\[
  D_\mu U=\partial_\mu U-i g\,a_\mu^aT^aU,
  \qquad
  \operatorname{tr}_{\square}(T^aT^b)=\delta^{ab}.
\]
The constants \(F_H>0\) and \(0<v_H\le1\) are low-energy coefficients of
the dense medium.  Symmetry fixes the tensor structure of
\eqref{eq:qcd-cfl-higgs-screening-action}, not the microscopic numerical
value of \(F_H\) or \(v_H\).
\end{controlledapproximation}

\paragraph{CFL Higgs quadratic mass matrix.}
In the pure-color CFL Higgs-screening effective action, before dynamical QED
mixing is included, the color-flavor-locked background gives equal quadratic
curvatures to all eight color gauge-field coordinates.  More precisely, within
Controlled Approximation~\ref{ca:qcd-cfl-higgs-screening-effective-action},
one has
\begin{equation}
  (M_{E,\rm quad}^2)^{ab}=g^2F_H^2\,\delta^{ab},
  \qquad
  (M_{M,\rm quad}^2)^{ab}=g^2F_H^2v_H^2\,\delta^{ab}.
  \label{eq:qcd-cfl-higgs-screening-masses}
\end{equation}
These are coefficients in a gauge-fixed quadratic effective action.  They are
not, by themselves, the gauge-invariant exponents
\(M_E^{\mathcal S},M_M^{\mathcal S}\) of
Definition~\ref{def:qcd-gauge-invariant-static-screening}.  Identifying a
particular renormalized Wilson-line field-strength exponent with \(gF_H\) or
\(gF_Hv_H\) requires a weak-coupling matching theorem for that operator scheme:
the line and endpoint renormalizations must be subtracted, the operator must
have nonzero overlap with the intended eigenchannel, and lower composite or
static-source states must be excluded or accounted for.

Use the gauge freedom to set \(U=1\) locally; this is unitary gauge for the
Higgs-coordinate chart, not a definition of a gauge-invariant operator.
Then
\[
  D_\mu U=-i g\,a_\mu^aT^a .
\]
Substitution into \eqref{eq:qcd-cfl-higgs-screening-action} gives
\[
  \mathcal L_{\rm H}^{(2)}
  =
  \frac{g^2F_H^2}{2}
  a_0^aa_0^b\operatorname{tr}(T^aT^b)
  -
  \frac{g^2F_H^2v_H^2}{2}
  a_i^aa_i^b\operatorname{tr}(T^aT^b).
\]
The trace convention of the chapter makes
\(\operatorname{tr}(T^aT^b)=\delta^{ab}\).  Therefore the quadratic
coefficients are exactly the matrices displayed in
\eqref{eq:qcd-cfl-higgs-screening-masses}.  The equality of the eight
eigenvalues follows from the unbroken diagonal \(SU(3)\): the gluons form
the adjoint representation of this diagonal group, and the quadratic mass
operator is an invariant bilinear form on the adjoint.  Since the adjoint of
\(SU(3)\) is simple, this invariant form is proportional to
\(\delta^{ab}\), and the direct computation fixes the proportionality
constant.

\begin{proposition}[Rotated electromagnetic response in ideal CFL]
\label{prop:qcd-cfl-rotated-electromagnetic-response}
Couple the ideal CFL effective coordinate to a dynamical electromagnetic
field \(A_\mu\) with coupling \(e\) and flavor generator
\[
  Q_{\rm em}
  =
  \operatorname{diag}\!\left(\frac23,-\frac13,-\frac13\right).
\]
In the same trace-delta convention as
Controlled Approximation~\ref{ca:qcd-cfl-higgs-screening-effective-action},
embed \(Q_{\rm em}\) in the color Cartan subalgebra and write
\[
  q_Q^2=\operatorname{tr}(Q_{\rm em}^2)=\frac23,
  \qquad
  \widehat Q=\frac{Q_{\rm em}}{q_Q},
  \qquad
  a_\mu^Q=\operatorname{tr}(a_\mu\widehat Q).
\]
The gauge-covariant derivative is
\begin{equation}
  D_\mu U
  =
  \partial_\mu U-i g\,a_\mu U+i e\,A_\mu UQ_{\rm em}.
  \label{eq:qcd-cfl-qed-covariant-derivative}
\end{equation}
In unitary gauge \(U=1\), the electric quadratic form in the
\((a_0^Q,A_0)\) sector is
\begin{equation}
  \frac{F_H^2}{2}
  \begin{pmatrix}a_0^Q&A_0\end{pmatrix}
  \begin{pmatrix}
    g^2 & -g e q_Q\\
    -g e q_Q & e^2 q_Q^2
  \end{pmatrix}
  \begin{pmatrix}a_0^Q\\ A_0\end{pmatrix}.
  \label{eq:qcd-cfl-rotated-em-mass-matrix}
\end{equation}
The determinant of this matrix is zero.  The massless long-range gauge field
is the rotated photon
\begin{equation}
  \widetilde A_\mu
  =
  \frac{g A_\mu+e q_Q a_\mu^Q}{\sqrt{g^2+e^2q_Q^2}},
  \label{eq:qcd-cfl-rotated-photon}
\end{equation}
while the orthogonal combination
\begin{equation}
  \widetilde G_\mu
  =
  \frac{g a_\mu^Q-e q_Q A_\mu}{\sqrt{g^2+e^2q_Q^2}}
  \label{eq:qcd-cfl-screened-em-gluon}
\end{equation}
has electric screening mass squared
\[
  m_{\widetilde G,E}^2
  =
  F_H^2(g^2+e^2q_Q^2).
\]
The spatial response has the same null and massive eigenvectors, with the
massive eigenvalue multiplied by \(v_H^2\).  The seven color directions
orthogonal to \(\widehat Q\) retain the CFL Higgs screening masses in
\eqref{eq:qcd-cfl-higgs-screening-masses}.

The inverse transformation is
\begin{equation}
  A_\mu
  =
  \frac{g\widetilde A_\mu-e q_Q\widetilde G_\mu}
       {\sqrt{g^2+e^2q_Q^2}},
  \qquad
  a_\mu^Q
  =
  \frac{e q_Q\widetilde A_\mu+g\widetilde G_\mu}
       {\sqrt{g^2+e^2q_Q^2}}.
  \label{eq:qcd-cfl-rotated-em-inverse}
\end{equation}
Thus the original color-Cartan coordinate \(a_\mu^Q\) is not a massive
eigenfield when \(e\ne0\).  In the diagonal quadratic approximation,
suppressing tensor projectors and contact terms,
\begin{equation}
  \langle a^Q a^Q\rangle
  =
  \frac{e^2q_Q^2}{g^2+e^2q_Q^2}
  \langle\widetilde A\,\widetilde A\rangle
  +
  \frac{g^2}{g^2+e^2q_Q^2}
  \langle\widetilde G\,\widetilde G\rangle .
  \label{eq:qcd-cfl-color-cartan-pole-content}
\end{equation}
It therefore contains a long-range rotated-photon contribution as well as a
massive rotated-gluon contribution.  Gauge-invariant observables that isolate
these eigenchannels must be built for the rotated combinations, for example
using the CFL orientation field or equivalent gauge-invariant composites,
rather than by identifying the bare color-Cartan field-strength line
correlator with a single quadratic eigenvalue.

Thus ideal CFL is not a superconductor for the unbroken rotated
\(U(1)\) gauge field.  Seven color directions orthogonal to \(\widehat Q\)
have massive quadratic response, the orthogonal photon-color combination
\(\widetilde G_\mu\) is massive, and the rotated photon
\(\widetilde A_\mu\) is long-ranged.
\end{proposition}

\begin{proof}
The right action in
\eqref{eq:qcd-cfl-qed-covariant-derivative} is the gauged version of the
locked flavor transformation \(U\mapsto hUg^{-1}\).  In the local chart
\(U=1\),
\[
  D_\mu U=-i g a_\mu+i e A_\mu Q_{\rm em}.
\]
Decompose the color gauge field as
\[
  a_\mu=a_\mu^\perp+a_\mu^Q\widehat Q,
  \qquad
  \operatorname{tr}(a_\mu^\perp\widehat Q)=0.
\]
Then
\[
  \operatorname{tr}\!\left[
    (g a_\mu-eA_\mu Q_{\rm em})^2
  \right]
  =
  g^2\operatorname{tr}\!\left[(a_\mu^\perp)^2\right]
  +(g a_\mu^Q-e q_QA_\mu)^2.
\]
This gives
\eqref{eq:qcd-cfl-rotated-em-mass-matrix} for the electric part of the
effective action and the corresponding factor of \(v_H^2\) for the spatial
part.  The two-by-two matrix has
\[
  \det
  \begin{pmatrix}
    g^2 & -g e q_Q\\
    -g e q_Q & e^2 q_Q^2
  \end{pmatrix}
  =
  g^2e^2q_Q^2-g^2e^2q_Q^2=0,
\]
and trace \(g^2+e^2q_Q^2\).  Its null vector is proportional to
\((e q_Q,g)\), giving \(\widetilde A_\mu\), and its nonzero eigenvector is
proportional to \((g,-e q_Q)\), giving \(\widetilde G_\mu\).  The remaining
color components are orthogonal to \(Q_{\rm em}\), so they do not mix with
the photon and keep the color-Higgs screening coefficients of
\eqref{eq:qcd-cfl-higgs-screening-masses}.
\end{proof}

\begin{definition}[CFL collective fields]
\label{def:qcd-cfl-collective-fields}
Ignoring quark-mass effects and the anomalous axial \(U(1)\), the
gauge-invariant low-energy collective fields of the CFL phase are:
\[
  \Sigma(x)\in SU(3),
  \qquad
  \Sigma\longmapsto g_L\Sigma g_R^{-1},
\]
for the chiral octet, and
\[
  \phi_B(x)\in U(1)_B/\mathbb Z_2,
  \qquad
  \delta\phi_B=\alpha_B
\]
locally for the baryon superfluid phase.  The quotient records the fact that
the gauge-invariant CFL order operators carrying baryon phase have even
baryon charge: a source-selected charge-two order parameter is unchanged by
the baryon rotation \(\alpha_B=\pi\).  Equivalently, one may use the
\(2\pi\)-periodic field \(\vartheta_B=2\phi_B\), which shifts by
\(2\alpha_B\).  The local derivative expansion below is insensitive to this
finite stabilizer, but large background gauge transformations are not.  The
leading real-time effective
Lagrangian in the rest frame of the dense medium is
\begin{equation}
  \mathcal L_{\rm CFL,NGB}
  =
  \frac{F_\pi^2}{4}
  \operatorname{tr}\!\left[
    \partial_0\Sigma\,\partial_0\Sigma^\dagger
    -
    v_\pi^2\partial_i\Sigma\,\partial_i\Sigma^\dagger
  \right]
  +
  \frac{F_B^2}{2}
  \left[
    (\partial_0\phi_B)^2
    -
    v_B^2(\partial_i\phi_B)^2
  \right]
  +\cdots .
  \label{eq:qcd-cfl-ngb-effective-action}
\end{equation}
The constants \(F_\pi,F_B,v_\pi,v_B\) are dense-medium low-energy
coefficients.  The ellipsis contains higher derivatives, mass-source terms,
and anomaly-suppressed effects.
\end{definition}

\paragraph{Physical collective-mode content of the ideal CFL coordinate chart.}
\label{par:qcd-cfl-collective-mode-content}
Under the assumptions encoded in
Definition~\ref{def:qcd-cfl-collective-fields}, the ideal CFL effective
coordinate chart contains
eight chiral Nambu--Goldstone modes and one baryon superfluid phonon.  To
quadratic order their dispersion relations are
\begin{equation}
  \omega^2=v_\pi^2\mathbf k^2
  \quad\text{for the chiral octet},
  \qquad
  \omega^2=v_B^2\mathbf k^2
  \quad\text{for the baryon phonon}.
  \label{eq:qcd-cfl-ngb-dispersions}
\end{equation}
The eight color-orientation modes are not additional physical
Nambu--Goldstone particles; in the screening effective action they are the
fields removed by the Higgs mechanism and are represented physically by the
eight screened color gauge sectors of the CFL Higgs-screening calculation
above.

The logical input is the effective coordinate chart; the independent dense-QCD
problem is the construction and control of that chart from a regulated QCD
definition.  The finite calculation is as follows.
Write
\[
  \Sigma(x)=\exp\!\left(\frac{2i\pi^a(x)\tau^a}{F_\pi}\right),
  \qquad
  \operatorname{tr}(\tau^a\tau^b)=\delta^{ab},
\]
where \(a=1,\ldots,8\).  Expanding
\eqref{eq:qcd-cfl-ngb-effective-action} to quadratic order gives
\[
  \mathcal L_{\pi}^{(2)}
  =
  \frac12\sum_{a=1}^8
  \left[
    (\partial_0\pi^a)^2
    -
    v_\pi^2(\partial_i\pi^a)^2
  \right],
\]
after the displayed normalization of the exponential is used.  The Euler
equation is
\[
  \partial_0^2\pi^a-v_\pi^2\nabla^2\pi^a=0,
\]
so a plane wave has \(\omega^2=v_\pi^2\mathbf k^2\).  The field
\(\Sigma\in SU(3)\) has dimension \(8\), hence the chiral sector contains
eight modes.

The baryon phase part of \eqref{eq:qcd-cfl-ngb-effective-action} gives
\[
  \partial_0^2\phi_B-v_B^2\nabla^2\phi_B=0,
\]
and therefore the second dispersion relation in
\eqref{eq:qcd-cfl-ngb-dispersions}.  This is one real compact phase field.
Finally, the color orientation \(U\) in
Controlled Approximation~\ref{ca:qcd-cfl-higgs-screening-effective-action}
has eight local coordinates, but it transforms under local color.  The
unitary-gauge computation in the CFL Higgs-screening calculation above shows
that those
coordinates are not gauge-invariant massless particles; they supply the
longitudinal screening components of the eight color gauge sectors.  Thus
the physical gapless count is \(8+1=9\), matching
Remark~\ref{rem:qcd-cfl-goldstone-count}.

\section{Anomaly Matching in the CFL Phase}
\label{sec:qcd-cfl-anomaly-matching}

The preceding CFL discussion identified gauge-invariant order diagnostics and
the physical collective fields.  A further consistency test is the response to
background fields for exact global symmetries.  This test is independent of
weak-coupling quasiparticle language: if the same microscopic QCD theory is
followed to long distances in a CFL phase, the background anomaly class of the
exact symmetry must be represented by the CFL infrared fields.

\begin{definition}[Faithful CFL symmetry and quotient backgrounds]
\label{def:qcd-cfl-faithful-global-backgrounds}
For massless \(N_c=N_f=3\) QCD, excluding the anomalous singlet axial
rotation, the connected continuous symmetry has Lie algebra
\[
  \mathfrak{su}(3)_L\oplus\mathfrak{su}(3)_R\oplus\mathfrak u(1)_B .
\]
The faithful connected group acting on local gauge-invariant operators is not
the direct product written at the Lie-algebra level.  With baryon number
normalized so that a color-singlet three-quark baryon has charge \(1\), it is
\begin{equation}
  G_{\rm QCD}^{0}
  =
  \frac{SU(3)_L\times SU(3)_R\times U(1)_B}
       {\mathbb Z_3^{V}},
  \qquad
  \mathbb Z_3^{V}
  =
  \{(\omega^k\mathbf 1,\omega^k\mathbf 1,1)\mid k=0,1,2\},
  \label{eq:qcd-cfl-faithful-connected-symmetry}
\end{equation}
where \(\omega=e^{2\pi i/3}\).  The diagonal flavor center is invisible to
mesonic operators and to the gauge-invariant baryon determinant, while a
left-only or right-only flavor center is visible on the chiral field
\(\Sigma\).

A background for \(G_{\rm QCD}^{0}\) is therefore not, in general, a triple of
independent \(SU(3)_L\), \(SU(3)_R\), and ordinary \(U(1)_B\) bundles.  On a
good cover, a quotient background can be represented by local lifts whose
flavor transition functions obey
\[
  g^L_{ij}g^L_{jk}g^L_{ki}
  =
  \omega^{n_{ijk}}\mathbf 1,
  \qquad
  g^R_{ij}g^R_{jk}g^R_{ki}
  =
  \omega^{n_{ijk}}\mathbf 1,
  \qquad
  n_{ijk}\in Z^2(M_4,\mathbb Z_3),
\]
with the same \(\mathbb Z_3\) cocycle for the left and right lifts.  The
fractional quark baryon charge is globalized by combining the dynamical color
bundle with the baryon line, not by choosing an independent cube root of a
baryon line.  Equivalently, the microscopic quark is coupled to a
color-baryon \(U(3)\) bundle \(\mathcal E_{cB}\) with local connection
\begin{equation}
  A_{cB}=a+\frac13B\,\mathbf 1_3,
  \qquad
  \det\mathcal E_{cB}=L_B ,
  \label{eq:qcd-cfl-color-baryon-u3-lift}
\end{equation}
where \(a\) is locally traceless and \(B\) is the determinant-line connection.
If one instead writes local \(SU(3)_c\) transitions and local cube roots of
the baryon transition functions, their two \(\mathbb Z_3\) triple-overlap
cocycles must cancel so that the quark bundle is globally defined.  A
\(2\pi\) baryon rotation of a quark is then a color-center phase, while the
same rotation is the identity on gauge-invariant integer-baryon operators.

All statements in this subsection are made on spin four-manifolds.  No
additional spin-charge quotient identifying \((-1)^F\) with a baryon rotation
is being imposed here; such a refinement would be extra background data.
\end{definition}

\begin{definition}[Lifted local CFL anomaly backgrounds]
\label{def:qcd-cfl-anomaly-backgrounds}
Work with massless \(N_f=N_c=3\) QCD and with the connected continuous
symmetry generated infinitesimally by
\[
  SU(3)_L\times SU(3)_R\times U(1)_B ,
\]
excluding the anomalous singlet axial rotation.  Choose a lift chart of the
faithful background of
Definition~\ref{def:qcd-cfl-faithful-global-backgrounds} in which the
\(\mathbb Z_3\) center cocycles vanish.  In that chart let \(A_L,A_R\) be
background \(SU(3)_L\) and \(SU(3)_R\) connections, and let \(B\) be the
connection on the baryon determinant line \(L_B\).  The local quark baryon
charge is
\[
  q_B(q)=\frac13 .
\]
This is the data needed to compute the de Rham, infinitesimal anomaly
polynomial.  It is not a complete specification of the anomaly theory on
arbitrary quotient bundles; the latter also includes the discrete center
cocycles, possible torsion response, and the large-gauge-transformation
normalization of the CFL Goldstone fields.
In this anomaly-matching subsection, flavor traces are taken in the
half-trace convention
\[
  \operatorname{tr}_{\rm f}(t^at^b)=\frac12\delta^{ab},
\]
the same convention used in the WZW level-matching argument leading to
\eqref{eq:wzw-level-forced-by-matching}.  This is only a convention for
anomaly traces; the trace-delta convention used elsewhere in this chapter is
related by \(\tau^a=\sqrt2\,t^a\).
\end{definition}

\begin{lemma}[UV lifted continuous flavor anomaly polynomial]
\label{lem:qcd-cfl-uv-anomaly-polynomial}
With the conventions of
Definition~\ref{def:qcd-cfl-anomaly-backgrounds}, the de Rham six-form
anomaly polynomial of the massless quarks in the lifted local chart, restricted
to \(SU(3)_L\times SU(3)_R\times U(1)_B\) curvature forms and omitting purely
gravitational terms, is
\begin{equation}
  I_6^{\rm UV}
  =
  \frac{N_c}{6(2\pi)^3}
  \left[
    \operatorname{tr}_{\rm f}(F_L^3)
    -
    \operatorname{tr}_{\rm f}(F_R^3)
  \right]
  +
  \frac{N_c q_B}{2(2\pi)^3}
  F_B\wedge
  \left[
    \operatorname{tr}_{\rm f}(F_L^2)
    -
    \operatorname{tr}_{\rm f}(F_R^2)
  \right],
  \label{eq:qcd-cfl-uv-anomaly-polynomial-general}
\end{equation}
where \(F_L,F_R,F_B\) are the corresponding curvatures.  For
\(N_c=3\) and \(q_B=1/3\), the mixed coefficient is
\begin{equation}
  N_cq_B/2=\frac12.
  \label{eq:qcd-cfl-mixed-anomaly-coefficient}
\end{equation}
\end{lemma}

\begin{proof}
The left-handed quark field consists of \(N_c\) color copies of a
left-handed Weyl fermion in the fundamental representation of \(SU(3)_L\)
with baryon charge \(q_B\).  For a single left-handed Weyl fermion in a
flavor bundle with connection \(A\), tensored with a \(U(1)\) line of charge
\(q_B\), the relevant degree-six Chern-character terms are
\begin{equation}
  \frac{1}{6(2\pi)^3}\operatorname{tr}_{\rm f}(F^3)
  +
  \frac{q_B}{2(2\pi)^3}F_B\wedge\operatorname{tr}_{\rm f}(F^2).
  \label{eq:qcd-cfl-single-weyl-anomaly-character}
\end{equation}
The terms proportional to
\(F_B^2\operatorname{tr}_{\rm f}F\) vanish because the flavor connection is
\(SU(3)\)-valued, and the pure \(F_B^3\) term cancels between the left- and
right-handed components of a vectorlike Dirac quark.  Multiplication by the
\(N_c\) color copies gives the left contribution.  A right-handed Weyl
fermion with the same representation and charge contributes with the
opposite sign to the four-dimensional chiral anomaly polynomial, giving the
minus signs in \eqref{eq:qcd-cfl-uv-anomaly-polynomial-general}.  Substituting
\(q_B=1/3\) gives \eqref{eq:qcd-cfl-mixed-anomaly-coefficient}.
\end{proof}

\begin{proposition}[CFL realization of the lifted continuous anomaly]
\label{prop:qcd-cfl-continuous-anomaly-matching}
Assume the ideal CFL infrared field content of
Definition~\ref{def:qcd-cfl-collective-fields}: a chiral field
\(\Sigma\in SU(3)\) with
\(\Sigma\mapsto g_L\Sigma g_R^{-1}\), and a local baryon-superfluid coordinate
\(\phi_B\) on \(U(1)_B/\mathbb Z_2\) which transforms infinitesimally under a
lifted \(U(1)_B\) background gauge transformation by
\begin{equation}
  B\mapsto B+d\lambda_B,
  \qquad
  \phi_B\mapsto\phi_B+\lambda_B .
  \label{eq:qcd-cfl-baryon-background-gauge-transformation}
\end{equation}
Then, in the lifted local chart of
Definition~\ref{def:qcd-cfl-anomaly-backgrounds}, the CFL infrared effective
action can represent the de Rham anomaly polynomial
\eqref{eq:qcd-cfl-uv-anomaly-polynomial-general} as follows.

First, the gauged Wess--Zumino--Witten functional of \(\Sigma\) has level
\begin{equation}
  n=N_c=3
  \label{eq:qcd-cfl-wzw-level}
\end{equation}
and matches the pure nonabelian chiral part
\begin{equation}
  \frac{N_c}{6(2\pi)^3}
  \left[
    \operatorname{tr}_{\rm f}(F_L^3)
    -
    \operatorname{tr}_{\rm f}(F_R^3)
  \right].
  \label{eq:qcd-cfl-pure-chiral-anomaly-piece}
\end{equation}
Second, define the invariant one-form
\begin{equation}
  D_B\phi_B=d\phi_B-B .
  \label{eq:qcd-cfl-baryon-goldstone-covariant-one-form}
\end{equation}
On a five-dimensional extension \(X_5\) of spacetime, the local inflow
representative
\begin{equation}
  \omega_{5,B}^{\rm CFL}
  =
  -
  \frac{D_B\phi_B}{2\pi}
  \wedge
  \frac{
    \operatorname{tr}_{\rm f}(F_L^2)
    -
    \operatorname{tr}_{\rm f}(F_R^2)
  }{2(2\pi)^2}
  \label{eq:qcd-cfl-baryon-mixed-anomaly-representative}
\end{equation}
satisfies
\begin{equation}
  d\omega_{5,B}^{\rm CFL}
  =
  \frac{F_B}{2\pi}
  \wedge
  \frac{
    \operatorname{tr}_{\rm f}(F_L^2)
    -
    \operatorname{tr}_{\rm f}(F_R^2)
  }{2(2\pi)^2}.
  \label{eq:qcd-cfl-baryon-mixed-anomaly-derivative}
\end{equation}
This is exactly the mixed \(U(1)_B\)-flavor term in
\eqref{eq:qcd-cfl-uv-anomaly-polynomial-general} for \(N_cq_B=1\).
On a general faithful quotient background this display is only the local
differential representative.  The exponentiated gauged WZW functional and the
baryon-Goldstone response must also be defined with the center cocycles of
Definition~\ref{def:qcd-cfl-faithful-global-backgrounds} and with the
\(\mathbb Z_2\) stabilizer of the charge-two CFL order parameter.  Matching
that torsion and large-gauge-transformation data is a stronger global anomaly
statement and is not claimed by this proposition.
\end{proposition}

\begin{proof}
The field \(\Sigma\) has the same transformation law as the ordinary chiral
field used in the WZW level-matching construction of
Chapter~\ref{ch:volume-ii-21}.  The anomaly-matching argument
uses only the number of microscopic color copies and the
\(SU(N_f)_L\) background anomaly coefficient; it does not use the
zero-density chiral condensate as the microscopic order parameter.  Therefore
the same level condition applies in the CFL phase:
\[
  n=N_c .
\]
For \(N_c=3\) this gives \eqref{eq:qcd-cfl-wzw-level}.  The gauged WZW
functional at this level has precisely the pure nonabelian chiral descent
class \eqref{eq:qcd-cfl-pure-chiral-anomaly-piece}.

It remains to check the terms involving the baryon background.  Under
\eqref{eq:qcd-cfl-baryon-background-gauge-transformation}, the combination
\(D_B\phi_B=d\phi_B-B\) is invariant.  Its exterior derivative is
\begin{equation}
  d(D_B\phi_B)=-F_B .
  \label{eq:qcd-cfl-d-dbphi}
\end{equation}
The Bianchi identities for the background flavor curvatures give
\[
  d\,\operatorname{tr}_{\rm f}(F_L^2)=0,
  \qquad
  d\,\operatorname{tr}_{\rm f}(F_R^2)=0 .
\]
Taking the exterior derivative of
\eqref{eq:qcd-cfl-baryon-mixed-anomaly-representative} and using
\eqref{eq:qcd-cfl-d-dbphi} gives
\eqref{eq:qcd-cfl-baryon-mixed-anomaly-derivative}.  Comparing with
\eqref{eq:qcd-cfl-uv-anomaly-polynomial-general}, and using \(N_cq_B=1\),
shows equality of the mixed anomaly polynomial.  The vector diagonal
subgroup is anomaly-free in both descriptions: setting \(F_L=F_R\) makes both
\eqref{eq:qcd-cfl-pure-chiral-anomaly-piece} and
\eqref{eq:qcd-cfl-baryon-mixed-anomaly-derivative} vanish.
\end{proof}

\begin{remark}[Goldstone normalization and local order parameters]
\label{rem:qcd-cfl-baryon-goldstone-normalization}
The scalar \(\phi_B\) in
\eqref{eq:qcd-cfl-baryon-background-gauge-transformation} is the local coset
angle shifted by the background \(U(1)_B\) gauge parameter.  It should not be
confused with the phase of a particular microscopic composite.  With
\(q_B(q)=1/3\), a local gauge-invariant determinant built from three diquark
factors has baryon charge \(2\), so the corresponding local order coordinate
is represented by a charge-two function, such as \(e^{2i\phi_B}\), in this
normalization.  The source-selected CFL order therefore breaks
\(U(1)_B\) to \(\mathbb Z_2\).  The anomaly calculation above uses the local
coupling of the compact Goldstone circle to the baryon background; the global
exponentiated response additionally remembers the \(\phi_B\sim\phi_B+\pi\)
identification and the quotient-background flux data.
\end{remark}

\section{Comparison of Confinement Criteria}

Several observables are often described by the same word ``confinement.''  In
this monograph they are kept distinct.

\begin{enumerate}
  \item The zero-temperature Wilson-loop area law is a statement about the
  large-loop asymptotics of a gauge-invariant loop operator and the static
  source potential.
  \item The thermal Polyakov-loop order parameter is a statement about
  center symmetry and the free energy of one static external charge.
  \item The 't Hooft loop probes magnetic or center-twist sectors and is
  dual, in favorable regimes, to Wilson-loop ordering.
  \item The center-vortex picture is a proposed infrared mechanism in which
  center-valued magnetic disorder controls the area law.
  \item The dual-superconductor picture is a proposed infrared mechanism in
  which condensation of magnetically charged objects squeezes electric flux.
\end{enumerate}

The first two items have clean operator definitions in lattice gauge theory.
The last three can be developed into sharp statements only after the
relevant line operators, disorder sectors, and continuum limits have been
specified.  Their value in this monograph is not as slogans, but as candidate
structures to be tested in examples and connected to the Wilsonian and BV
definitions of gauge theory developed earlier.

This comparison is the endpoint of the chapter's phase-specification
discipline.  A Polyakov loop, a Wilson loop, a diquark diagnostic, a
transport coefficient, and a susceptibility are not interchangeable signals:
they live in different limiting procedures and respond to different sources.
When two such signals are correlated in a regime of QCD, that correlation is
a dynamical conclusion to be proved, simulated with controlled errors, or
recorded as a conjectural continuation.  It is not part of the definition of
the phase.

\begin{openproblem}[Continuum QCD phase diagram with controlled errors]
\label{op:qcd-phase-diagram-controlled-errors}
Construct the continuum QCD phase diagram in the \((T,\mu_B,m)\) variables
with mathematically controlled thermodynamic and continuum limits, including
the order of the pure-Yang--Mills deconfinement transition, the chiral
transition in massless limits, the physical crossover region, the existence
or nonexistence of a critical endpoint, and the color-superconducting
regions at high baryon density.
\end{openproblem}
