\chapter{Hydrodynamic Fluctuations And Long-Time Tails}
\label{ch:thermal-hydrodynamic-fluctuations}

\ConventionsRealTime

Hydrodynamics describes the long-wavelength evolution of conserved
densities.  In a thermal state those densities also fluctuate.  The
fluctuations are fixed at the linear level by static susceptibilities and
the KMS fluctuation--dissipation relation; their nonlinear couplings then
produce nonanalytic low-frequency terms in real-time correlators.  These
terms are the hydrodynamic long-time tails.  They are universal within the
hydrodynamic effective theory because their nonanalytic part is determined by
the conserved-mode propagators and by the local nonlinear vertices.

Throughout this chapter spacetime is \(\mathbb R_t\times\mathbb R^d\), with
\(d\) spatial dimensions.  Fourier modes are
\(\exp(-\ii\omega t+\ii\mathbf k\cdot\mathbf x)\).  The diffusion
normalizations are those of
Chapter~\ref{ch:thermal-sk-hydrodynamic-effective-actions}: the charge
susceptibility is \(\chi=\partial n/\partial\mu\), the conductivity is
\(\sigma\), and \(D=\sigma/\chi\).  Equal-time thermal fluctuations of a
coarse-grained density carry the covariance \(T\chi\).

\section{Static Fluctuation Data}

Let \(q_A(t,\mathbf x)\) be conserved densities in a homogeneous KMS state
at temperature \(T\) and chemical potentials \(\mu_A\).  Write
\[
  \delta q_A=q_A-\omega_\beta(q_A),
\]
where \(\omega_\beta\) denotes the thermal state.  The static susceptibility
matrix in the chemical-potential convention is
\[
  \chi_{AB}
  =
  \left(\frac{\partial q_A}{\partial\mu_B}\right)_T .
\]
For hydrodynamic variables averaged over a cell whose size is large compared
with microscopic correlation lengths, the equal-time connected covariance is
\begin{equation}
  \int d^dx\,
  \omega_\beta\!\left(
    \delta q_A(0,\mathbf x)\delta q_B(0,\mathbf0)
  \right)_{\rm conn}
  =
  T\chi_{AB}.
  \label{eq:hydro-fluct-static-covariance}
\end{equation}
Equivalently, if one uses the dimensionless thermodynamic force
\(\alpha_A=\mu_A/T\), the covariance equals
\(\partial q_A/\partial\alpha_B\).  This convention removes a common source
of factor-of-\(T\) errors in stochastic hydrodynamics.

For one diffusive charge density \(n\), the Gaussian equilibrium
probability functional for long-wavelength configurations is
\begin{equation}
  P_{\rm eq}[n]\propto
  \exp\!\left[
    -\frac{1}{2T\chi}\int d^dx\,(\delta n)^2
  \right],
  \qquad
  \delta n=n-\omega_\beta(n).
  \label{eq:hydro-fluct-static-gaussian}
\end{equation}
This formula is a local central-limit statement inside the hydrodynamic
window.  The continuum delta function in
\eqref{eq:hydro-fluct-static-covariance} is an effective-theory notation for
a cell-averaged covariance; the cell scale is kept between the microscopic
correlation length and the external wavelength.

\section{Linear Stochastic Diffusion}

The first-order stochastic constitutive relation for one conserved charge is
\begin{equation}
  \partial_t n+\partial_i j^i=0,
  \qquad
  j^i=-D\partial_i n+\xi^i,
  \label{eq:hydro-fluct-langevin-diffusion}
\end{equation}
with Gaussian white current noise
\begin{equation}
  \langle\xi^i(t,\mathbf x)\xi^j(t',\mathbf x')\rangle
  =
  2T\sigma\,\delta^{ij}
  \delta(t-t')\delta^{(d)}(\mathbf x-\mathbf x').
  \label{eq:hydro-fluct-current-noise}
\end{equation}
The coefficient \(2T\sigma\) is the same coefficient derived from the
dynamical KMS symmetry of the diffusion action in
Chapter~\ref{ch:thermal-sk-hydrodynamic-effective-actions}.  The stochastic
equation gives
\[
  n(\omega,\mathbf k)
  =
  \frac{\ii k_i\xi^i(\omega,\mathbf k)}
       {-\ii\omega+Dk^2},
  \qquad
  k^2=\mathbf k\cdot\mathbf k .
\]
Hence the symmetrized density correlator is
\begin{equation}
  G^{\rm sym}_{nn}(\omega,\mathbf k)
  =
  \frac{2T\sigma k^2}
       {\omega^2+(Dk^2)^2}.
  \label{eq:hydro-fluct-density-symmetric}
\end{equation}
Integrating over frequency gives
\[
  \int\frac{d\omega}{2\pi}
  G^{\rm sym}_{nn}(\omega,\mathbf k)
  =
  \frac{T\sigma}{D}
  =
  T\chi,
\]
which is exactly the static covariance
\eqref{eq:hydro-fluct-static-covariance}.  Thus the Einstein relation
\(D=\sigma/\chi\) is forced by compatibility of the deterministic diffusion
constant, the conductivity, and the equilibrium fluctuation measure.

The retarded density source-response kernel from the Schwinger--Keldysh
action is
\begin{equation}
  K^R_{nn}(\omega,\mathbf k)
  =
  \chi\,\frac{Dk^2}{Dk^2-\ii\omega}.
  \label{eq:hydro-fluct-diffusive-retarded}
\end{equation}
Its imaginary part obeys the low-frequency fluctuation--dissipation relation
\[
  G^{\rm sym}_{nn}(\omega,\mathbf k)
  =
  \frac{2T}{\omega}\,
  \operatorname{Im}K^R_{nn}(\omega,\mathbf k)
  +O(\omega/T)
\]
inside the classical hydrodynamic regime.  The exact quantum formula is the
KMS relation of Chapter~\ref{ch:thermal-spectral-kubo-transport}; the
displayed expression is its \(|\omega|\ll T\) limit.

\section{Local Schwinger--Keldysh Origin}

The stochastic equation \eqref{eq:hydro-fluct-langevin-diffusion} is the
Hubbard--Stratonovich form of the local diffusion action
\begin{equation}
  I_{\rm diff}
  =
  \int dt\,d^dx\,
  \left[
  \chi B_{a0}B_{r0}
  -\sigma B_{ai}\partial_tB_{ri}
  +\ii T\sigma B_{ai}B_{ai}
  \right].
  \label{eq:hydro-fluct-diffusion-sk-action}
\end{equation}
Here \(B_{r\mu}=A_{r\mu}+\partial_\mu\varphi_r\) and
\(B_{a\mu}=A_{a\mu}+\partial_\mu\varphi_a\) are the gauge-invariant
hydrodynamic source combinations.  The \(a\)-phase variation imposes the
continuity equation, and the imaginary term has the Gaussian representation
\[
  \exp\!\left[
    -T\sigma\int B_{ai}B_{ai}
  \right]
  =
  \int D\xi\,
  \exp\!\left[
    -\int\frac{\xi_i\xi_i}{4T\sigma}
    +\ii\int B_{ai}\xi_i
  \right].
\]
The same coefficient therefore controls current response, current noise, and
diffusive relaxation.  This is the local effective-action statement of KMS:
one chooses the dissipative term and the noise term together.

\section{Finite-Cell Current Fluctuations and the Macroscopic Cost}
\label{sec:hydro-fluct-finite-current-mft}

The white current noise in
\eqref{eq:hydro-fluct-current-noise} can be obtained from a finite
empirical-flow calculation before any continuum stochastic field notation is
introduced.  This is the finite origin of the Gaussian current cost used in
macroscopic fluctuation theory.  The continuum functional is a hydrodynamic
scaling statement; the exact statement at this stage is the contraction of a
finite jump-flow relative entropy.

Consider one oriented mesoscopic bond.  During a time block, let \(q_+\) and
\(q_-\) be the empirical rates of jumps across the bond in the positive and
negative directions.  At fixed local occupation data the corresponding
reference jump intensities are positive numbers \(r_+\) and \(r_-\).  The
local contribution to the finite level-\(2.5\) cost of
Section~\ref{sec:finite-empirical-flow-level-25} is
\begin{equation}
  \Phi(q_+,q_-;r_+,r_-)
  =
  q_+\log\frac{q_+}{r_+}-q_++r_+
  +
  q_-\log\frac{q_-}{r_-}-q_-+r_- ,
  \label{eq:hydro-fluct-finite-bond-flow-cost}
\end{equation}
with \(q_\pm\ge0\).  The net empirical current through the bond is
\[
  j=q_+-q_- .
\]
For fixed \(j\), strict convexity gives a unique minimizer.  Introducing a
Lagrange multiplier \(\ell\) for \(q_+-q_-=j\) gives
\[
  \log\frac{q_+}{r_+}+\ell=0,
  \qquad
  \log\frac{q_-}{r_-}-\ell=0,
\]
hence
\[
  q_+q_-=r_+r_-,
  \qquad
  q_+-q_-=j .
\]
Writing
\[
  \Delta(j)=\sqrt{j^2+4r_+r_-},
\]
the minimizer is
\begin{equation}
  q_+^*(j)=\frac{\Delta(j)+j}{2},
  \qquad
  q_-^*(j)=\frac{\Delta(j)-j}{2}.
  \label{eq:hydro-fluct-bond-current-minimizer}
\end{equation}
Substitution into \eqref{eq:hydro-fluct-finite-bond-flow-cost} gives the
exact finite current cost
\begin{equation}
  \Phi_{\rm cur}(j;r_+,r_-)
  =
  j\log
  \frac{j+\sqrt{j^2+4r_+r_-}}{2r_+}
  -
  \sqrt{j^2+4r_+r_-}
  +r_++r_- .
  \label{eq:hydro-fluct-finite-bond-current-cost}
\end{equation}
It vanishes at the typical current
\[
  j_0=r_+-r_-,
  \qquad
  \Delta(j_0)=r_++r_- .
\]
The derivative of the contracted cost is the multiplier conjugate to the
current,
\[
  \frac{d\Phi_{\rm cur}}{dj}
  =
  \log\frac{q_+^*(j)}{r_+},
\]
and therefore
\[
  \left.
  \frac{d^2\Phi_{\rm cur}}{dj^2}
  \right|_{j=j_0}
  =
  \frac{1}{r_++r_-}.
\]
The current cost near its minimum is consequently
\begin{equation}
  \Phi_{\rm cur}(j;r_+,r_-)
  =
  \frac{(j-j_0)^2}{2(r_++r_-)}
  +O\!\left((j-j_0)^3\right).
  \label{eq:hydro-fluct-current-cost-quadratic-expansion}
\end{equation}
The same exact expression also records the entropy-production asymmetry:
\begin{equation}
  \Phi_{\rm cur}(j;r_+,r_-)
  -
  \Phi_{\rm cur}(-j;r_+,r_-)
  =
  -j\log\frac{r_+}{r_-}.
  \label{eq:hydro-fluct-bond-current-gc-asymmetry}
\end{equation}
Thus the symmetric traffic \(r_++r_-\) fixes the Gaussian current variance,
while the ratio \(r_+/r_-\) fixes the antisymmetric entropy-production
part.  No continuum limit has entered this calculation.

To pass from the finite current cost to hydrodynamics, one must add scaling
input.  Suppose a sequence of finite cell regulators has a local-equilibrium
diffusive scaling limit in which the conserved density is \(n(t,\mathbf x)\)
and the empirical current is \(j^i(t,\mathbf x)\).  Suppose moreover that
the finite typical current and finite traffic converge in the local
hydrodynamic window to the deterministic current and covariance density.  In
a local basis where the positive symmetric conductivity matrix is diagonal,
this componentwise statement is
\[
  j^i_{\rm det}(n)
  =
  -D^{ij}(n)\partial_jn+O(\partial^2),
  \qquad
  r_+^i+r_-^i
  \longrightarrow
  2T\,\sigma^{ij}(n)
\]
in the normalization of \eqref{eq:hydro-fluct-current-noise}.  Then the
quadratic Taylor term
\eqref{eq:hydro-fluct-current-cost-quadratic-expansion} becomes the local
Gaussian current cost
\begin{equation}
  \mathcal I_{\rm cur}[n,j]
  =
  \frac{1}{4T}
  \int dt\,d^dx\,
  \bigl(j^i+D^{ik}\partial_kn\bigr)
  (\sigma^{-1})_{ij}
  \bigl(j^j+D^{j\ell}\partial_\ell n\bigr),
  \label{eq:hydro-fluct-mft-gaussian-current-cost}
\end{equation}
subject to the continuity equation
\begin{equation}
  \partial_t n+\partial_i j^i=0 .
  \label{eq:hydro-fluct-mft-continuity-constraint}
\end{equation}
Equation \eqref{eq:hydro-fluct-mft-gaussian-current-cost} is the
Onsager--Machlup current form of the stochastic diffusion theory
\eqref{eq:hydro-fluct-langevin-diffusion}.  Integrating over \(j\) with the
constraint \eqref{eq:hydro-fluct-mft-continuity-constraint}, or equivalently
introducing a Lagrange multiplier for the constraint, gives the same
quadratic Schwinger--Keldysh action
\eqref{eq:hydro-fluct-diffusion-sk-action} after the
Hubbard--Stratonovich step above.  The non-Gaussian terms in the exact
finite expression \eqref{eq:hydro-fluct-finite-bond-current-cost} are
irrelevant in the central-limit hydrodynamic fluctuation regime but are
essential for far-from-typical current large deviations when the underlying
finite regulator is retained.

The logical status is therefore sharp.  The finite contraction
\eqref{eq:hydro-fluct-finite-bond-current-cost} is a direct finite
consequence
of the regulated jump-flow rate function.  The macroscopic functional
\eqref{eq:hydro-fluct-mft-gaussian-current-cost} is obtained from it only
after one assumes, or proves in a specified model, the local-equilibrium
hydrodynamic scaling limit, the convergence of traffic to \(2T\sigma\), and
the Einstein relation between \(D\), \(\sigma\), and \(\chi\).

\section{Hydrodynamic Nonlinearities}

Hydrodynamic variables are local coordinates on the manifold of equilibrium
states.  Expanding the currents in fluctuations around a homogeneous state
gives
\begin{equation}
  J^i_A
  =
  J^i_{A,\rm lin}
  +\frac12\,\lambda^i{}_{A,BC}\,\phi_B\phi_C
  +O(\phi^3,\phi\,\partial\phi,\partial^2\phi),
  \label{eq:hydro-fluct-quadratic-current}
\end{equation}
where \(\phi_A=\delta q_A\) after diagonalizing the linear hydrodynamic
modes, and \(\lambda^i{}_{A,BC}\) is symmetric in \(B,C\).  The coefficient
\(\lambda\) is determined by the equation of state and by the chosen
hydrodynamic frame.  For example, the convective part of the stress tensor
contains \(\pi^i\pi^j/(\varepsilon+p)\), where \(\pi^i=T^{0i}\) is momentum
density and \(\varepsilon+p\) is the enthalpy density.

The quadratic term in \eqref{eq:hydro-fluct-quadratic-current} produces a
loop of hydrodynamic modes in current-current correlators.  The calculation
is an effective-theory calculation with an ultraviolet cutoff
\(\Lambda_{\rm hyd}\) below the inverse microscopic length.  Local analytic
terms depend on this cutoff and renormalize local transport coefficients.
The low-frequency nonanalytic part is independent of the detailed cutoff
profile.

\section{Scalar Equilibrium Tail From A Diffusive Pair}

Consider one scalar diffusive mode \(\phi\) with
\[
  \langle\phi(t,\mathbf k)\phi(0,-\mathbf k)\rangle
  =
  T\chi\,e^{-Dk^2t},
  \qquad t>0.
\]
Let the scalar composite operator be
\[
  O(t,\mathbf x)=\lambda:\!\phi(t,\mathbf x)^2\!: .
\]
This operator is a scalar in a homogeneous isotropic parity-even state.  A
spatial current needs a vector supplied by a gradient, momentum density, or an
external anisotropy.  Thus the calculation below is the scalar-composite
equilibrium tail; it becomes a current or stress calculation only after the
corresponding tensor vertex is supplied.

At Gaussian order, the connected classical or symmetrized equilibrium
correlator
\[
  C^{\rm sym}_{OO}(t)
  =
  \int d^dx\,
  \langle O(t,\mathbf x)O(0,\mathbf0)\rangle_{\rm conn}^{\rm sym}
\]
has the Wick contraction
\begin{align}
  C^{\rm sym}_{OO}(t)
  &=
  2\lambda^2(T\chi)^2
  \int\frac{d^dk}{(2\pi)^d}\,e^{-2Dk^2t}
  \nonumber\\
  &=
  2\lambda^2(T\chi)^2\,
  (8\pi Dt)^{-d/2},
  \qquad t>0 .
  \label{eq:hydro-fluct-time-tail-single-diffusive-mode}
\end{align}
This is the basic long-time tail in an equilibrium fluctuation channel.  More
general hydrodynamic systems give sums of terms of the same form, with
diffusion constants replaced by sums of the damping rates of the two modes in
the loop and with tensor projectors from momentum or sound modes.  The
exponent is fixed by the Gaussian integration over small loop momentum.

\section{Symmetric, Retarded, And Transport Nonanalyticities}

The frequency-domain scalar object obtained directly from
\eqref{eq:hydro-fluct-time-tail-single-diffusive-mode} is the symmetrized
equilibrium spectrum.  Its one-sided Laplace kernel is the pair-diffusion
integral
\begin{equation}
  I_d(z;D)
  =
  \int\frac{d^dk}{(2\pi)^d}
  \frac{1}{z+2Dk^2},
  \qquad
  z=-\ii\omega+0^+ .
  \label{eq:hydro-fluct-basic-loop-integral}
\end{equation}
The analytic terms are cutoff dependent.  The nonanalytic part is obtained
by subtracting local Taylor terms, or equivalently by dimensional
regularization of the infrared integral:
\begin{equation}
  I_d^{\rm nonan}(z;D)
  =
  \frac{\Gamma(1-d/2)}{(4\pi)^{d/2}(2D)^{d/2}}\,
  z^{d/2-1},
  \qquad d\notin 2\mathbb Z_{\ge1}.
  \label{eq:hydro-fluct-loop-nonanalytic-general}
\end{equation}
For the two-sided symmetrized spectrum the real completion of this boundary
value gives
\begin{equation}
  G^{\rm sym}_{OO}(\omega)\big|_{\rm nonan}
  \propto
  |\omega|^{d/2-1}
  \qquad
  (d\notin2\mathbb Z_{\ge1}),
  \label{eq:hydro-fluct-symmetric-spectrum-exponent}
\end{equation}
with the logarithmic replacement in even spatial dimension.  This is the
Fourier transform of the \(t^{-d/2}\) equilibrium tail.

The retarded correlator is a different object.  For the commutator convention
of Chapter~\ref{ch:thermal-spectral-kubo-transport},
\[
  G^{\rm sym}_{OO}(\omega)
  =
  \frac12\coth\!\left(\frac{\omega}{2T}\right)\rho_{OO}(\omega),
  \qquad
  \rho_{OO}=-2\,\operatorname{Im}G^R_{OO}.
\]
In the hydrodynamic regime,
\begin{equation}
  \rho_{OO}(\omega)\big|_{\rm nonan}
  =
  \frac{\omega}{T}\,
  G^{\rm sym}_{OO}(\omega)\big|_{\rm nonan}
  +O(\omega^3/T^3),
  \label{eq:hydro-fluct-fdt-extra-frequency}
\end{equation}
so the spectral density, and hence the nonlocal retarded boundary value, has
one additional power of frequency:
\begin{equation}
  G^R_{OO}(\omega)\big|_{\rm nonan}
  =
  c_R\,(-\ii\omega+0^+)^{d/2}
  \quad
  \text{up to local real polynomials}
  \label{eq:hydro-fluct-retarded-exponent-shift}
\end{equation}
for non-even \(d\), with \(c_R\) fixed by the vertex and by the retarded
completion.  Thus in \(d=3\) the scalar equilibrium spectrum has a
\(|\omega|^{1/2}\) cusp, while the retarded scalar correlator has a
\((- \ii\omega)^{3/2}\) branch term.  In \(d=2\), the pair is
\begin{equation}
  G^{\rm sym}_{OO}\big|_{\rm nonan}\propto
  \log\frac{\Lambda_{\rm hyd}^2D}{|\omega|},
  \qquad
  G^R_{OO}\big|_{\rm nonan}\propto
  (-\ii\omega)\log\frac{\Lambda_{\rm hyd}^2D}{-\ii\omega+0^+}.
  \label{eq:hydro-fluct-d2-fdt-log-pair}
\end{equation}
The integral \(I_d\) therefore has two legitimate later uses: it can describe
the symmetrized scalar spectrum, and after a Kubo tensor structure divides a
retarded correlator by \(-\ii\omega\), it can describe the nonanalytic
transport kernel.  By itself it lacks the vector or tensor vertex and the FDT
frequency factor required for a retarded current or stress correlator.

\paragraph{Infrared loop integral by Schwinger parameter.}
For \(\operatorname{Re}z>0\) and \(D>0\), the dimensionally regularized
infrared integral
\begin{equation}
  \int\frac{d^dk}{(2\pi)^d}\frac{1}{z+2Dk^2}
  \label{eq:hydro-fluct-dimreg-loop-start}
\end{equation}
is
\begin{equation}
  \frac{\Gamma(1-d/2)}{(4\pi)^{d/2}(2D)^{d/2}}z^{d/2-1}.
  \label{eq:hydro-fluct-dimreg-loop-derived}
\end{equation}
The causal boundary value used in retarded kernels is obtained by analytic
continuation to
\begin{equation}
  z=-\ii\omega+0^+,
  \label{eq:hydro-fluct-retarded-z-branch}
\end{equation}
with the principal logarithm, so that the result is analytic for
\(\operatorname{Im}\omega>0\).

For \(\operatorname{Re}z>0\),
\begin{equation}
  \frac{1}{z+2Dk^2}
  =
  \int_0^\infty ds\,e^{-s(z+2Dk^2)} .
  \label{eq:hydro-fluct-schwinger-representation}
\end{equation}
The Gaussian integral in \(d\) dimensions is
\begin{equation}
  \int\frac{d^dk}{(2\pi)^d}e^{-2Dsk^2}
  =
  (4\pi\,2Ds)^{-d/2}.
  \label{eq:hydro-fluct-gaussian-d-dimensional}
\end{equation}
Therefore
\begin{equation}
  I_d(z;D)
  =
  \frac{1}{(4\pi)^{d/2}(2D)^{d/2}}
  \int_0^\infty ds\,s^{-d/2}e^{-zs}.
  \label{eq:hydro-fluct-schwinger-after-gaussian}
\end{equation}
The remaining integral is the defining analytic continuation of the gamma
function:
\begin{equation}
  \int_0^\infty ds\,s^{\alpha-1}e^{-zs}
  =
  \Gamma(\alpha)z^{-\alpha},
  \qquad \alpha=1-\frac d2 .
  \label{eq:hydro-fluct-gamma-integral}
\end{equation}
Substituting \(\alpha=1-d/2\) gives
\eqref{eq:hydro-fluct-dimreg-loop-derived}.  Dimensional regularization
drops power-divergent local terms and retains the nonanalytic infrared
dependence.  The causal prescription
\eqref{eq:hydro-fluct-retarded-z-branch} follows from the original
denominator \(z+2Dk^2=-\ii\omega+0^+ +2Dk^2\), which is nonzero for
\(\operatorname{Im}\omega>0\).

In even spatial dimension the pole of \(\Gamma(1-d/2)\) is replaced, after
local subtraction, by \(z^{d/2-1}\log z\).  The first cases are
\[
\begin{array}{c|c}
d & I_d^{\rm nonan}(z;D)\\ \hline
1 & \displaystyle \frac{1}{2(2D)^{1/2}}\,z^{-1/2}\\[2mm]
2 & \displaystyle -\frac{1}{8\pi D}\,\log z\\[2mm]
3 & \displaystyle -\frac{1}{4\pi(2D)^{3/2}}\,z^{1/2}
\end{array}
\]
up to scheme-dependent additive constants in \(d=2\).  The branch is fixed
by \(z=-\ii\omega+0^+\), so the corresponding causal kernel is analytic in
the upper half \(\omega\)-plane.

For \(d=3\), an explicit cutoff computation displays the separation between
local and nonlocal terms:
\[
  I_3^\Lambda(z;D)
  =
  \frac{1}{2\pi^2(2D)}
  \left[
    \Lambda
    -
    \left(\frac{z}{2D}\right)^{1/2}
    \arctan\!\left(
      \frac{\Lambda}{(z/2D)^{1/2}}
    \right)
  \right].
\]
As \(z/\Lambda^2\to0\),
\[
  I_3^\Lambda(z;D)
  =
  \frac{\Lambda}{4\pi^2D}
  -
  \frac{z^{1/2}}{4\pi(2D)^{3/2}}
  +O(z/\Lambda).
\]
The first term renormalizes a local transport coefficient.  The square-root
term is the symmetrized scalar-spectrum nonanalyticity, or the transport-kernel
nonanalyticity after the retarded correlator has been divided by its Kubo
factor.  The retarded correlator itself carries the additional frequency
factor in \eqref{eq:hydro-fluct-fdt-extra-frequency}.

\section{Transport Coefficients As Scale-Dependent Local Data}

Let \(\kappa\) be a transport coefficient multiplying a local term in the
hydrodynamic constitutive relation or Schwinger--Keldysh action.  Hydrodynamic
loops with momenta below \(\Lambda_{\rm hyd}\) shift the coefficient:
\[
  \kappa_{\rm phys}(\omega)
  =
  \kappa(\Lambda_{\rm hyd})
  +
  \Delta\kappa_{\rm loc}(\Lambda_{\rm hyd})
  +
  \Delta\kappa_{\rm nonan}(\omega).
\]
The sum of the first two terms is the local coefficient appropriate to the
chosen hydrodynamic cutoff.  The last term is fixed by hydrodynamic mode
propagation and is a genuinely nonlocal-in-time low-frequency contribution,
distinct from a finite local derivative coefficient.  This is the
finite-temperature analogue of ordinary effective-field theory
renormalization: local coefficients are scheme dependent, while nonanalytic
infrared dependence is physical within the regime of validity.

The order of limits is part of the definition.  One first takes the
thermodynamic limit of the KMS state, then probes frequencies and momenta
inside the hydrodynamic scaling regime, and only then studies the small
\(\omega,\mathbf k\) asymptotics.  Interchanging this order can erase the
hydrodynamic pole, mix static and dynamic susceptibilities, or replace
retarded functions by finite-volume quasiperiodic sums.

If the operator channel defines a transport kernel through
\[
  G^R_{\mathcal T\mathcal T}(\omega,\mathbf0)
  =
  G^R_{\mathcal T\mathcal T}(0,\mathbf0)
  -\ii\omega\,\kappa(\omega)
  +\cdots ,
\]
then a retarded nonanalyticity
\[
  \delta G^R_{\mathcal T\mathcal T}(\omega)
  \propto
  (-\ii\omega)^{d/2}
\]
corresponds to
\begin{equation}
  \delta\kappa(\omega)
  =
  \frac{\delta G^R_{\mathcal T\mathcal T}(\omega)}{-\ii\omega}
  \propto
  (-\ii\omega)^{d/2-1}.
  \label{eq:hydro-fluct-transport-kernel-exponent}
\end{equation}
This is the location where the \(I_d\) exponent \(d/2-1\) enters a transport
coefficient.  In \(d=3\), the retarded correlator has a \(3/2\)-power branch
while the frequency-dependent viscosity or conductivity kernel has a
square-root correction.  In \(d=2\), the transport kernel runs logarithmically
with frequency; a fixed finite bare coefficient is then a scale-dependent local
parameter of the hydrodynamic effective theory, not a frequency-independent
infrared observable.

\section{Stress-Tensor Noise}

For a relativistic fluid in the local rest frame, write the stochastic
spatial stress as
\[
  T^{ij}
  =
  p\,\delta^{ij}
  -\eta
  \left(
    \partial^iv^j+\partial^jv^i
    -\frac{2}{d}\delta^{ij}\partial_kv^k
  \right)
  -\zeta\,\delta^{ij}\partial_kv^k
  +\Xi^{ij}
  +O(v^2,\partial^2).
\]
The white-noise covariance required by local KMS is
\begin{align}
  \langle\Xi^{ij}(x)\Xi^{kl}(x')\rangle
  &=
  2T\Bigl[
    \eta
    \left(
      \delta^{ik}\delta^{jl}
      +\delta^{il}\delta^{jk}
      -\frac{2}{d}\delta^{ij}\delta^{kl}
    \right)
    +\zeta\,\delta^{ij}\delta^{kl}
  \Bigr]
  \nonumber\\
  &\qquad\qquad\times
  \delta(t-t')\delta^{(d)}(\mathbf x-\mathbf x').
  \label{eq:hydro-fluct-stress-noise}
\end{align}
For any real symmetric test tensor \(s_{ij}\), decompose
\[
  s_{ij}=s^{\rm T}_{ij}+\frac{1}{d}\delta_{ij}s^k{}_k,
  \qquad
  \delta^{ij}s^{\rm T}_{ij}=0.
\]
Contracting the tensor in \eqref{eq:hydro-fluct-stress-noise} with
\(s_{ij}s_{kl}\) gives
\[
  4T\eta\,s^{\rm T}_{ij}s^{\rm T}_{ij}
  +
  2T\zeta\,(s^k{}_k)^2 .
\]
Thus positivity of the stochastic stress covariance is equivalent, at this
order, to \(\eta\ge0\) and \(\zeta\ge0\).  The same inequalities are obtained
from nonnegative entropy production in deterministic hydrodynamics.

\section{Momentum Tails}

Momentum density \(\pi^i=T^{0i}\) has static covariance
\[
  \langle\pi^i(\mathbf k)\pi^j(-\mathbf k)\rangle_{\rm eq}
  =
  T(\varepsilon+p)\delta^{ij}
\]
in the local rest frame.  Transverse momentum diffuses with kinematic
viscosity
\[
  \nu=\frac{\eta}{\varepsilon+p}.
\]
Because the stress tensor contains the convective term
\[
  T^{ij}_{\rm conv}
  =
  \frac{\pi^i\pi^j}{\varepsilon+p},
\]
the shear-stress correlator contains a two-momentum hydrodynamic loop.  Its
late-time behavior is again of the form
\[
  \langle T^{ij}(t)T^{kl}(0)\rangle_{\rm conn}
  \sim
  T^2\,\mathcal P^{ij,kl}\,
  (8\pi\nu t)^{-d/2},
\]
with \(\mathcal P^{ij,kl}\) determined by the transverse projectors and by
angular averaging.  Sound modes add oscillatory pieces with damping
controlled by the sound attenuation constant.  The existence of the tail
therefore follows from the convective nonlinearity and the conserved
momentum pole; the exact tensor coefficient requires the full hydrodynamic
mode basis and frame convention.

\paragraph{A symmetry-allowed \(d=3\) shear loop.}
The scalar \(\phi^2\) example above becomes a genuine shear-stress example
when the tensor vertex is
\[
  T^{xy}_{\rm conv}
  =
  \frac{\pi^x\pi^y}{w},
  \qquad
  w=\varepsilon+p .
\]
For the transverse shear modes,
\[
  \langle\pi^i(t,\mathbf k)\pi^j(0,-\mathbf k)\rangle_{\rm T}
  =
  T w\,P_{\rm T}^{ij}(\hat{\mathbf k})\,e^{-\nu k^2t},
  \qquad
  P_{\rm T}^{ij}=\delta^{ij}-\hat k^i\hat k^j .
\]
Wick contraction gives
\[
  C^{\rm sym}_{xy,xy;{\rm T}}(t)
  =
  T^2\int\frac{d^3k}{(2\pi)^3}
  \left[
    P_{\rm T}^{xx}P_{\rm T}^{yy}
    +
    P_{\rm T}^{xy}P_{\rm T}^{yx}
  \right]e^{-2\nu k^2t}.
\]
The angular average is
\begin{equation}
  \left\langle
    P_{\rm T}^{xx}P_{\rm T}^{yy}
    +
    P_{\rm T}^{xy}P_{\rm T}^{yx}
  \right\rangle_{S^2}
  =
  1-\frac13-\frac13+2\frac1{15}
  =
  \frac7{15}.
  \label{eq:hydro-fluct-transverse-angular-seven-fifteenths}
\end{equation}
Thus
\begin{equation}
  C^{\rm sym}_{xy,xy;{\rm T}}(t)
  =
  \frac7{15}\,T^2(8\pi\nu t)^{-3/2}.
  \label{eq:hydro-fluct-shear-transverse-tail}
\end{equation}

The retarded SK loop has one retarded momentum propagator and one symmetric
momentum propagator.  With a sharp hydrodynamic cutoff, its transverse
nonlocal part can be written
\begin{equation}
  \delta G^R_{xy,xy;{\rm T}}(\omega)
  =
  -\frac{7T}{30\pi^2}
  \int_0^{\Lambda_{\rm hyd}}dk\,
  \frac{k^4}{k^2-\ii\omega/(2\nu)} .
  \label{eq:hydro-fluct-shear-retarded-radial-integral}
\end{equation}
The radial identity
\[
  \int_0^\Lambda dk\,\frac{k^4}{k^2+a}
  =
  \frac{\Lambda^3}{3}-a\Lambda+a^{3/2}
  \arctan\frac{\Lambda}{a^{1/2}}
\]
with \(a=-\ii\omega/(2\nu)\) gives
\begin{align}
  \delta G^R_{xy,xy;{\rm T}}(\omega)
  &=
  -\frac{7T}{90\pi^2}\Lambda_{\rm hyd}^3
  -\ii\omega\,\frac{7T\Lambda_{\rm hyd}}{60\pi^2\nu}
  \nonumber\\
  &\quad
  -\frac{7T}{60\pi(2\nu)^{3/2}}\,
  (-\ii\omega+0^+)^{3/2}
  +O(\omega^2/\Lambda_{\rm hyd}) .
  \label{eq:hydro-fluct-shear-retarded-three-halves}
\end{align}
The cubic term shifts the pressure contact, the linear term shifts the local
viscosity at the cutoff, and the branch term is the retarded long-time-tail
contribution.  Dividing the branch term by \(-\ii\omega\) in the shear Kubo
structure gives the frequency-dependent viscosity correction
\[
  \delta\eta_{\rm T}(\omega)
  \propto
  (-\ii\omega+0^+)^{1/2}.
\]

Sound modes use the longitudinal projector
\[
  P_{\rm L}^{ij}=\hat k^i\hat k^j,
  \qquad
  \left\langle
    P_{\rm L}^{xx}P_{\rm L}^{yy}
    +
    P_{\rm L}^{xy}P_{\rm L}^{yx}
  \right\rangle_{S^2}
  =
  \frac2{15}.
\]
Their propagating poles carry denominators
\(\omega\mp c_s k+\ii\Gamma_s k^2/2\).  After the frequency convolution, the
same infrared radial structure as
\eqref{eq:hydro-fluct-shear-retarded-radial-integral} appears with the sound
attenuation replacing the shear diffusion constant and with the longitudinal
projector coefficient above.  The sound-sound loop therefore contributes the
same retarded branch power \((-\ii\omega)^{3/2}\), while mixed
shear-sound pieces are analytic at this order because their poles sit on
different hydrodynamic scalings.  The exponent is universal; the coefficient
records the full mode content and frame convention.

\section{Microscopic Theorem Boundary}

A microscopic derivation from QFT must specify the KMS state, conserved
currents, thermodynamic susceptibilities, the thermodynamic limit, clustering
estimates, and the topology in which hydrodynamic correlators emerge from
real-time QFT correlators.  It must also identify the slow subspace of
conserved densities, prove local equilibration in the scaling regime being
used, and derive the Schwinger--Keldysh positivity and KMS constraints on
the effective action.  Hydrodynamic fluctuation theory then computes the
consequences of these inputs.

The long-time-tail calculation is rigorous as an internal statement of the
stochastic hydrodynamic effective theory with a stated cutoff and Gaussian
leading measure.  Promoting it to a theorem about a particular interacting
continuum QFT requires a proof that the QFT has the hydrodynamic scaling
limit assumed above.

\begin{openproblem}[Hydrodynamic limit]
\label{op:hydrodynamic-limit-qft}
Derive fluctuating hydrodynamics and long-time tails directly from an
interacting QFT with a KMS state, including the hydrodynamic scaling limit,
noise kernel, nonlinear couplings, transport-coefficient renormalization,
and control of the order of thermodynamic, real-time, and low-frequency
limits.
\end{openproblem}
