\chapter{Hydrodynamics from Ward Identities}
\label{ch:thermal-hydrodynamics-from-ward-identities}

\ConventionsLorentzian

Hydrodynamics is the long-wavelength effective theory of conserved
densities.  Its variables are local thermodynamic parameters attached to a
family of states whose correlators are controlled, at long distances and long
times, by Ward identities and by an expansion in gradients.  This chapter
develops the
deterministic first-order theory in flat spacetime and fixes the sign and
normalization conventions that connect it to the Kubo formulae of
Chapter~\ref{ch:thermal-spectral-kubo-transport}.

Let \(D=d+1\), where \(d\) is the number of spatial dimensions.  The
Minkowski metric is mostly plus, \(\eta_{\mu\nu}=(-,+,\ldots,+)\).  Repeated
ordinary charge labels \(A,B\) are summed.  All formulas in this chapter are
written for parity-even fluids without anomalies; anomalous transport is
developed separately in Chapter~\ref{ch:thermal-anomalous-topological-transport}.

\section{Conserved Densities and Local State Variables}

Let \(T^{\mu\nu}\) be the stress tensor and let \(J^\mu_A\) be conserved
currents for ordinary internal symmetries.  In flat space without external
sources,
\begin{equation}
  \partial_\mu T^{\mu\nu}=0,
  \qquad
  \partial_\mu J^\mu_A=0 .
  \label{eq:hydro-flat-ward-identities}
\end{equation}
A local equilibrium state is parametrized by a future-timelike velocity field
\(u^\mu(x)\), temperature \(T(x)>0\), and chemical potentials
\(\mu_A(x)\), with
\[
  u^\mu u_\mu=-1.
\]
The spatial projector in the local rest frame is
\[
  \Delta^{\mu\nu}=\eta^{\mu\nu}+u^\mu u^\nu,
  \qquad
  \Delta^{\mu\nu}u_\nu=0.
\]

The ideal constitutive relations are
\begin{equation}
  T^{\mu\nu}_{(0)}
  =
  \varepsilon u^\mu u^\nu
  +p\,\Delta^{\mu\nu},
  \qquad
  J^\mu_{A(0)}=n_Au^\mu .
  \label{eq:hydro-ideal-constitutive-relations}
\end{equation}
Here \(\varepsilon(T,\mu)\), \(p(T,\mu)\), and \(n_A(T,\mu)\) are local
thermodynamic functions of the underlying QFT.  The frame used below is the
Landau frame:
\begin{equation}
  u_\mu T^{\mu\nu}=-\varepsilon u^\nu,
  \label{eq:hydro-landau-frame}
\end{equation}
so the nonideal stress tensor carries no energy flux in the local rest
frame.  Charge diffusion is allowed to have components orthogonal to
\(u^\mu\).

\section{Thermodynamic Closure}

The equation of state supplies a function \(p=p(T,\mu_A)\).  Define
\[
  s=\left(\frac{\partial p}{\partial T}\right)_{\mu},
  \qquad
  n_A=\left(\frac{\partial p}{\partial\mu_A}\right)_{T,\mu_{B\ne A}},
\]
and
\begin{equation}
  \varepsilon+p=Ts+\mu_A n_A.
  \label{eq:hydro-gibbs-duhem-energy}
\end{equation}
Then
\begin{equation}
  dp=s\,dT+n_A\,d\mu_A,
  \qquad
  d\varepsilon=T\,ds+\mu_A\,dn_A .
  \label{eq:hydro-thermodynamic-relations}
\end{equation}
These identities are the local thermodynamic closure of the ideal Ward
system.  They are imposed pointwise on the slowly varying local state
variables; the microscopic QFT input is the equilibrium equation of state and
the assumption that the hydrodynamic scaling limit probes only this local
state manifold.

\section{Ideal Equations and Entropy Conservation}

\begin{proposition}[Ideal hydrodynamic equations]
\label{prop:ideal-hydro-equations}
Substituting \eqref{eq:hydro-ideal-constitutive-relations} into the Ward
identities gives
\begin{align}
  u^\mu\partial_\mu\varepsilon
  +(\varepsilon+p)\partial_\mu u^\mu&=0,
  \label{eq:hydro-ideal-energy-equation}\\
  (\varepsilon+p)u^\mu\partial_\mu u^\alpha
  +\Delta^{\alpha\mu}\partial_\mu p&=0,
  \label{eq:hydro-ideal-euler-equation}\\
  u^\mu\partial_\mu n_A+n_A\partial_\mu u^\mu&=0 .
  \label{eq:hydro-ideal-charge-equation}
\end{align}
\end{proposition}

\begin{proof}
Contract \(\partial_\mu T^{\mu\nu}_{(0)}=0\) with \(u_\nu\).  Since
\[
  u_\nu\partial_\mu(\varepsilon u^\mu u^\nu)
  =
  -\partial_\mu(\varepsilon u^\mu),
  \qquad
  u_\nu\partial_\mu(p\Delta^{\mu\nu})
  =
  -p\,\partial_\mu u^\mu,
\]
using \(u_\nu\partial_\mu u^\nu=\frac12\partial_\mu(u_\nu u^\nu)=0\), one
obtains \eqref{eq:hydro-ideal-energy-equation}.  Projecting the same Ward
identity with \(\Delta^\alpha{}_\nu\) gives
\[
  \Delta^\alpha{}_\nu\partial_\mu(\varepsilon u^\mu u^\nu)
  =
  \varepsilon u^\mu\partial_\mu u^\alpha,
  \qquad
  \Delta^\alpha{}_\nu\partial_\mu(p\Delta^{\mu\nu})
  =
  p\,u^\mu\partial_\mu u^\alpha+\Delta^{\alpha\mu}\partial_\mu p,
\]
which is \eqref{eq:hydro-ideal-euler-equation}.  Current conservation gives
\eqref{eq:hydro-ideal-charge-equation} directly.
\end{proof}

\begin{proposition}[Ideal entropy current]
\label{prop:ideal-entropy-conservation}
The ideal entropy current
\[
  S^\mu_{(0)}=s u^\mu
\]
is conserved on ideal hydrodynamic solutions:
\[
  \partial_\mu S^\mu_{(0)}=0.
\]
\end{proposition}

\begin{proof}
Using the first law,
\[
  T u^\mu\partial_\mu s
  =
  u^\mu\partial_\mu\varepsilon
  -\mu_A u^\mu\partial_\mu n_A .
\]
Substitute the ideal energy equation
\eqref{eq:hydro-ideal-energy-equation} and the ideal charge equations
\eqref{eq:hydro-ideal-charge-equation}.  This gives
\[
  T u^\mu\partial_\mu s
  =
  -(\varepsilon+p)\partial_\mu u^\mu
  +\mu_A n_A\partial_\mu u^\mu
  =
  -Ts\,\partial_\mu u^\mu,
\]
where \(\varepsilon+p=Ts+\mu_A n_A\) was used.  Therefore
\[
  \partial_\mu(su^\mu)
  =
  u^\mu\partial_\mu s+s\partial_\mu u^\mu=0.
\]
\end{proof}

\section{First-Order Constitutive Relations}

Define the expansion, acceleration, and shear tensor by
\[
  \vartheta=\partial_\mu u^\mu,
  \qquad
  a^\mu=u^\nu\partial_\nu u^\mu,
\]
and
\begin{equation}
  \sigma^{\mu\nu}
  =
  2\Delta^{\mu\alpha}\Delta^{\nu\beta}
  \left(
  \partial_{(\alpha}u_{\beta)}
  -\frac{1}{d}\Delta_{\alpha\beta}\vartheta
  \right).
  \label{eq:hydro-shear-tensor-definition}
\end{equation}
Then \(\sigma^{\mu\nu}u_\nu=0\), \(\sigma^\mu{}_\mu=0\), and
\(\sigma^{\mu\nu}=\sigma^{\nu\mu}\).

In Landau frame, the parity-even first-order constitutive relations are
\begin{align}
  T^{\mu\nu}
  &=
  \varepsilon u^\mu u^\nu
  +p\Delta^{\mu\nu}
  -\eta\,\sigma^{\mu\nu}
  -\zeta\,\Delta^{\mu\nu}\vartheta
  +O(\partial^2),
  \label{eq:hydro-first-order-stress}\\
  J_A^\mu
  &=
  n_Au^\mu
  -T\,\Sigma_{AB}\Delta^{\mu\nu}
  \partial_\nu\!\left(\frac{\mu_B}{T}\right)
  +O(\partial^2).
  \label{eq:hydro-first-order-current}
\end{align}
The matrix \(\Sigma_{AB}\) is the conductivity matrix in the normalization
where, at constant \(T\), Fick diffusion has diffusion matrix
\[
  D_{AB}=\Sigma_{AC}(\chi^{-1})_{CB},
  \qquad
  \chi_{AB}=\left(\frac{\partial n_A}{\partial\mu_B}\right)_T .
\]

\section{Entropy Production at First Order}

The first-order entropy current in Landau frame is
\begin{equation}
  S^\mu
  =
  s u^\mu-\frac{\mu_A}{T}\nu_A^\mu+O(\partial^2),
  \qquad
  \nu_A^\mu
  =
  -T\Sigma_{AB}\Delta^{\mu\nu}
  \partial_\nu\!\left(\frac{\mu_B}{T}\right).
  \label{eq:hydro-first-order-entropy-current}
\end{equation}

\begin{proposition}[Entropy production fixes positivity]
\label{prop:first-order-entropy-production}
Using \eqref{eq:hydro-first-order-stress} and
\eqref{eq:hydro-first-order-current}, the entropy production through second
order in gradients is
\begin{equation}
  \partial_\mu S^\mu
  =
  \frac{\eta}{2T}\sigma^{\mu\nu}\sigma_{\mu\nu}
  +\frac{\zeta}{T}\vartheta^2
  +T\,\Sigma_{AB}
  \Delta^{\mu\nu}
  \partial_\mu\!\left(\frac{\mu_A}{T}\right)
  \partial_\nu\!\left(\frac{\mu_B}{T}\right)
  +O(\partial^3).
  \label{eq:hydro-entropy-production}
\end{equation}
Hence local entropy production is nonnegative for arbitrary first gradients
if
\[
  \eta\ge0,\qquad \zeta\ge0,\qquad
  \Sigma_{AB}\ \text{is positive semidefinite}.
\]
\end{proposition}

\begin{proof}
Write the dissipative stress tensor as
\[
  \tau^{\mu\nu}
  =
  -\eta\sigma^{\mu\nu}-\zeta\Delta^{\mu\nu}\vartheta,
  \qquad
  T^{\mu\nu}=T_{(0)}^{\mu\nu}+\tau^{\mu\nu}.
\]
Using the Ward identities, the thermodynamic identities, and
\(u_\mu\tau^{\mu\nu}=u_\mu\nu_A^\mu=0\), the divergence of
\eqref{eq:hydro-first-order-entropy-current} through second order reduces to
\[
  \partial_\mu S^\mu
  =
  -\frac{1}{T}\tau^{\mu\nu}\partial_\mu u_\nu
  -\nu_A^\mu\partial_\mu\!\left(\frac{\mu_A}{T}\right)
  +O(\partial^3).
\]
The decomposition of \(\partial_\mu u_\nu\) into shear, trace, vorticity,
and acceleration gives
\[
  \sigma^{\mu\nu}\partial_\mu u_\nu
  =
  \frac12\sigma^{\mu\nu}\sigma_{\mu\nu},
  \qquad
  \Delta^{\mu\nu}\partial_\mu u_\nu=\vartheta.
\]
Substituting \(\tau^{\mu\nu}\) and \(\nu_A^\mu\) gives
\eqref{eq:hydro-entropy-production}.
\end{proof}

\section{Charge Diffusion}

Consider a state with \(u^\mu=(1,\mathbf0)\), constant \(T\), and vanishing
background flow.  Linearize a single conserved charge around equilibrium.
The constitutive relation gives
\[
  J^0=n_0+\delta n,\qquad
  J^i=-\Sigma\,\partial_i\delta\mu,
  \qquad
  \delta n=\chi\,\delta\mu .
\]
Charge conservation gives
\begin{equation}
  \partial_t\delta n-D\nabla^2\delta n=0,
  \qquad
  D=\frac{\Sigma}{\chi}.
  \label{eq:hydro-charge-diffusion-equation}
\end{equation}
For a plane wave \(\delta n\propto e^{-\ii\omega t+\ii\mathbf k\cdot\mathbf x}\),
the pole is
\begin{equation}
  \omega=-\ii D k^2+O(k^4).
  \label{eq:hydro-charge-diffusion-pole}
\end{equation}
The Einstein relation \(D=\Sigma/\chi\) is the hydrodynamic form of the same
coefficient matching that appears in the Kubo conductivity formula.

\section{Neutral Linearized Modes}

Linearize a neutral fluid around a homogeneous rest state:
\[
  u^\mu=(1,\mathbf v)+O(\mathbf v^2),
  \qquad
  \varepsilon=\varepsilon_0+\delta\varepsilon,
  \qquad
  p=p_0+\delta p,
  \qquad
  \delta p=c_s^2\delta\varepsilon ,
\]
where
\[
  c_s^2=\left(\frac{\partial p}{\partial\varepsilon}\right)_{\rm eq},
  \qquad
  w=\varepsilon_0+p_0 .
\]
The stress tensor components to first order in fluctuations and first order
in gradients are
\begin{align}
  T^{00}&=\varepsilon_0+\delta\varepsilon,
  \label{eq:hydro-linear-t00}\\
  T^{0i}&=wv^i,
  \label{eq:hydro-linear-t0i}\\
  T^{ij}
  &=
  p_0\delta^{ij}
  +c_s^2\delta\varepsilon\,\delta^{ij}
  -\eta\left(\partial_i v_j+\partial_j v_i
  -\frac{2}{d}\delta_{ij}\partial_\ell v^\ell\right)
  -\zeta\delta_{ij}\partial_\ell v^\ell .
  \label{eq:hydro-linear-tij}
\end{align}
The Ward identities become
\begin{align}
  \partial_t\delta\varepsilon+w\partial_i v^i&=0,
  \label{eq:hydro-linear-energy}\\
  w\partial_t v_i+\partial_i\delta p
  -\eta\nabla^2v_i
  -\left(\zeta+\eta\left(1-\frac{2}{d}\right)\right)
  \partial_i\partial_jv^j&=0 .
  \label{eq:hydro-linear-momentum}
\end{align}

\begin{proposition}[Shear and sound poles]
\label{prop:hydro-shear-sound-poles}
For fluctuations proportional to
\(e^{-\ii\omega t+\ii\mathbf k\cdot\mathbf x}\), the transverse velocity
mode has
\begin{equation}
  \omega_{\rm shear}
  =
  -\ii\,\frac{\eta}{w}k^2+O(k^4).
  \label{eq:hydro-shear-pole}
\end{equation}
The longitudinal energy-momentum modes obey
\begin{equation}
  \omega_{\rm sound,\pm}
  =
  \pm c_s k
  -\frac{\ii}{2}\Gamma_s k^2
  +O(k^3),
  \qquad
  \Gamma_s
  =
  \frac{\zeta+2\eta\frac{d-1}{d}}{w}.
  \label{eq:hydro-sound-poles}
\end{equation}
\end{proposition}

\begin{proof}
For a transverse mode, \(k_i v^i=0\) and
\eqref{eq:hydro-linear-momentum} becomes
\[
  -\ii\omega w v_i+\eta k^2v_i=0,
\]
which gives \eqref{eq:hydro-shear-pole}.

For a longitudinal mode take \(\mathbf k=k\hat x\) and \(v_i=v_L\hat x_i\).
Equations \eqref{eq:hydro-linear-energy} and
\eqref{eq:hydro-linear-momentum} give
\[
  -\ii\omega\delta\varepsilon+\ii wk v_L=0,
\]
and
\[
  -\ii\omega wv_L+\ii k c_s^2\delta\varepsilon
  +\left(\zeta+2\eta\frac{d-1}{d}\right)k^2v_L=0 .
\]
Eliminating \(v_L\) gives
\[
  \omega^2-c_s^2k^2
  +\ii\frac{\zeta+2\eta(d-1)/d}{w}\,\omega k^2=0.
\]
Solving perturbatively at small \(k\) gives
\eqref{eq:hydro-sound-poles}.
\end{proof}

\section{Retarded Correlator Pole Matching}

Hydrodynamic poles are singularities of retarded correlators after the
thermodynamic limit and the hydrodynamic scaling limit have been taken in the
specified order.  For a single diffusive charge with susceptibility \(\chi\),
the density response to a source for the chemical potential has the universal
long-wavelength form
\begin{equation}
  G^R_{nn}(\omega,\mathbf k)
  =
  \chi\,
  \frac{Dk^2}{Dk^2-\ii\omega}
  +\text{terms analytic at }\omega,k=0 .
  \label{eq:hydro-density-retarded-diffusion}
\end{equation}
in the source convention where \(G^R_{nn}(0,\mathbf k\to0)=\chi\).  The pole
is \(\omega=-\ii Dk^2\).  Taking \(\mathbf k\to0\) before
\(\omega\to0\) removes the diffusive denominator; taking the hydrodynamic
scaling limit \(\omega\sim k^2\) keeps it.  This is the same order-of-limits
issue that appears in dc conductivity and static susceptibility.

In the shear channel, the momentum-density correlator has a pole at
\eqref{eq:hydro-shear-pole}.  In the sound channel, the stress-tensor and
energy-density correlators have the pair of poles
\eqref{eq:hydro-sound-poles}.  The residues are fixed by thermodynamic
susceptibilities and Ward identities, while the imaginary parts of the pole
locations are fixed by the Kubo coefficients \(\eta\), \(\zeta\), and
\(\Sigma_{AB}\).

\section{Hydrodynamic Limit as a QFT Claim}

The deterministic equations above are consequences of a proposed
hydrodynamic scaling limit of the microscopic QFT.  The limit has the
following structure.  First take the thermodynamic limit of the KMS state.
Then consider sources and correlators with
\[
  \omega\tau_{\rm micro}\ll1,
  \qquad
  |\mathbf k|\ell_{\rm micro}\ll1,
\]
while keeping the ratios appropriate to the mode under study:
\(\omega\sim k^2\) for diffusion and shear, and \(\omega\sim k\) for sound.
The QFT assertion is that all nonconserved excitations produce analytic
contributions in this scaling region, and the nonanalyticities are the
conservation-law poles derived above.  Proving this assertion requires
clustering, real-time analyticity, local equilibration, and control of
long-time tails beyond the deterministic approximation.

\begin{openproblem}[Hydrodynamic emergence theorem]
\label{op:hydrodynamic-emergence-theorem}
Find physically useful QFT hypotheses under which local thermalization,
clustering, analyticity of retarded functions, positivity of entropy
production, and the hydrodynamic pole expansion can be proved from the
microscopic theory.
\end{openproblem}
