\chapter{Schwinger--Keldysh Hydrodynamic Effective Actions}
\label{ch:thermal-sk-hydrodynamic-effective-actions}

\ConventionsRealTime

Hydrodynamic Schwinger--Keldysh effective actions are real-time doubled
generating functionals for conserved densities in a long-wavelength state.
Their defining constraints are inherited from the closed time path:
normalization at coincident sources, a reality relation, positivity of the
imaginary part, gauge or diffeomorphism Ward identities, and the KMS condition
when the initial state is thermal.  This chapter develops the quadratic
multi-charge diffusion action explicitly, because it is the simplest place
where all of these constraints can be seen without hiding the
source-response signs or the matrix structure of susceptibilities and
conductivities.

Let spacetime be \(\mathbb R_t\times\mathbb R^d\).  Spatial indices
\(i,j\) are raised with \(\delta^{ij}\), and
\(\nabla^2=\partial_i\partial_i\).  Fourier modes are written as
\(\exp(-\ii\omega t+\ii\mathbf k\cdot\mathbf x)\) in the derivations below.
The microscopic retarded commutator convention is the one fixed in
Chapter~\ref{ch:thermal-spectral-kubo-transport}.  In this chapter
\(K^R\) denotes the source-response kernel obtained by differentiating the
Schwinger--Keldysh functional with respect to the \(a\)-type source.  The
conversion between \(K^R\) and the commutator retarded function includes the
sign convention of the microscopic source coupling and possible local contact
terms.

\section{Closed-Time-Path Constraints}

Let \(A_{1\mu}\) and \(A_{2\mu}\) be background gauge fields on the two
branches of the Schwinger--Keldysh contour.  For an initial density operator
\(\rho_0\), define
\begin{equation}
  Z[A_1,A_2]
  =
  \operatorname{Tr}\!\left(
  U[A_1]\rho_0U[A_2]^\dagger
  \right).
  \label{eq:sk-hydro-generating-functional}
\end{equation}
Introduce average and difference variables
\begin{equation}
  A_{r\mu}=\frac12(A_{1\mu}+A_{2\mu}),
  \qquad
  A_{a\mu}=A_{1\mu}-A_{2\mu}.
  \label{eq:sk-hydro-ra-sources}
\end{equation}
If \(Z=\exp(\ii I_{\rm SK})\), unitarity and Hermiticity imply
\begin{align}
  I_{\rm SK}[A_r,A_a=0]&=0,
  \label{eq:sk-hydro-unitarity-zero}\\
  I_{\rm SK}[A_r,A_a]^*
  &=
  -I_{\rm SK}[A_r,-A_a],
  \label{eq:sk-hydro-reality}\\
  \operatorname{Im} I_{\rm SK}[A_r,A_a]&\ge0
  \label{eq:sk-hydro-positivity}
\end{align}
for source configurations in the domain of the effective description.
Equation \eqref{eq:sk-hydro-unitarity-zero} says that every term in the
effective action contains at least one \(a\)-type field.  The positivity
condition is the damping condition in the path-integral weight
\(\exp(\ii I_{\rm SK})\).

\begin{proposition}[Finite closed-time-path constraints]
\label{prop:finite-closed-time-path-constraints}
Suppose first that the Hilbert space is finite-dimensional, that
\(\rho_0\ge0\) is a density matrix with \(\operatorname{Tr}\rho_0=1\), and
that the branch evolutions \(U[A_s]\) are unitary.  Then
\[
  Z[A,A]=1,\qquad
  Z[A_1,A_2]^*=Z[A_2,A_1],\qquad
  |Z[A_1,A_2]|\le1 .
  \label{eq:finite-ctp-constraints}
\]
Consequently, on the branch of \(I_{\rm SK}=-\ii\log Z\) continuous from
\(I_{\rm SK}[A_r,0]=0\), equations
\eqref{eq:sk-hydro-unitarity-zero}--\eqref{eq:sk-hydro-positivity} follow.
\end{proposition}

\begin{proof}
The diagonal identity is
\[
  Z[A,A]=\operatorname{Tr}(U[A]\rho_0U[A]^\dagger)
  =\operatorname{Tr}\rho_0=1 .
\]
Complex conjugation gives
\[
  Z[A_1,A_2]^*
  =
  \operatorname{Tr}\!\left(
  U[A_2]\rho_0U[A_1]^\dagger
  \right)=Z[A_2,A_1],
  \label{eq:finite-ctp-reality-proof}
\]
where cyclicity of the finite trace and \(\rho_0^\dagger=\rho_0\) are used.
For the bound, introduce the Hilbert--Schmidt inner product
\((X,Y)_{\rm HS}=\operatorname{Tr}(X^\dagger Y)\) and set
\[
  X=U[A_2]\rho_0^{1/2},\qquad
  Y=U[A_1]\rho_0^{1/2}.
\]
Then \(Z[A_1,A_2]=(X,Y)_{\rm HS}\).  Cauchy--Schwarz gives
\[
  |Z[A_1,A_2]|^2
  \le
  \operatorname{Tr}(\rho_0)\operatorname{Tr}(\rho_0)=1 .
  \label{eq:finite-ctp-cauchy-schwarz}
\]
Since \(Z=\exp(\ii I_{\rm SK})\), the last inequality is
\(\exp(-\operatorname{Im}I_{\rm SK})\le1\), hence
\(\operatorname{Im}I_{\rm SK}\ge0\).  The reality relation for
\(I_{\rm SK}\) follows from \eqref{eq:finite-ctp-reality-proof} after
rewriting branch exchange as \(A_a\mapsto -A_a\) at fixed \(A_r\).
\end{proof}

For continuum QFT the same algebra is first applied with an ultraviolet
regulator and finite volume, and only then passed to the thermodynamic and
continuum limits.  Without this order of limits the inequality
\(|Z|\le1\) is a statement about the regulated closed-time-path functional,
not a theorem about an unconstructed real-time path integral.

\section{Hydrodynamic Phase Field and Gauge Invariance}

For \(N\) commuting conserved \(U(1)\) charges, labelled by
\(A,B=1,\ldots,N\), the hydrodynamic modes are represented by phase fields
\[
  \varphi^A_1,\qquad \varphi^A_2 .
\]
The branchwise gauge transformations are
\[
  A^A_{s\mu}\mapsto A^A_{s\mu}+\partial_\mu\lambda^A_s,
  \qquad
  \varphi^A_s\mapsto\varphi^A_s-\lambda^A_s,
  \qquad s=1,2.
\]
Thus the gauge-invariant combinations are
\begin{equation}
  B^A_{s\mu}=A^A_{s\mu}+\partial_\mu\varphi^A_s .
  \label{eq:sk-hydro-branch-b-fields}
\end{equation}
In \(r/a\) variables,
\begin{equation}
  B^A_{r\mu}=A^A_{r\mu}+\partial_\mu\varphi^A_r,
  \qquad
  B^A_{a\mu}=A^A_{a\mu}+\partial_\mu\varphi^A_a .
  \label{eq:sk-hydro-ra-b-fields}
\end{equation}
The field \(\varphi^A_a\) is the Lagrange-multiplier partner of the
conservation law for the \(A\)-th charge.  Varying the action with respect
to \(\varphi^A_a\) is equivalent to the corresponding Ward identity because
\(\delta B^A_{a\mu}=\partial_\mu\delta\varphi^A_a\).

\begin{remark}[Gauge invariance and the hydrodynamic Ward identity]
\label{rem:sk-hydro-gauge-invariance-ward-identity}
Let \(I_{\rm eff}\) be a local functional of the gauge-invariant variables
\(B^A_{r\mu},B^A_{a\mu}\).  Define the expectation-value current on the
physical slice \(A_a=\varphi_a=0\) by
\[
  J_A^\mu(x)
  =
  \left.
  \frac{\delta I_{\rm eff}}{\delta A^A_{a\mu}(x)}
  \right|_{A_a=\varphi_a=0}.
  \label{eq:sk-hydro-current-functional-derivative}
\]
Then the Euler--Lagrange equation for \(\varphi^A_a\) is the distributional
Ward identity
\[
  \partial_\mu J_A^\mu=0
  \label{eq:sk-hydro-current-ward-identity}
\]
in the absence of external anomaly terms.  Equivalently, the \(a\)-phase is
not an additional physical scalar; it is the doubled-field implementation of
current conservation.

On the physical slice, \(B_{a\mu}^A=A_{a\mu}^A+\partial_\mu\varphi_a^A\).
Thus the variation of the action under a compactly supported
\(\delta\varphi_a^A\) is
\[
  \delta I_{\rm eff}
  =
  \int dt\,d^dx\,
  J_A^\mu\,\partial_\mu\delta\varphi_a^A
  =
  -\int dt\,d^dx\,
  (\partial_\mu J_A^\mu)\delta\varphi_a^A ,
  \label{eq:sk-hydro-ward-variation-proof}
\]
with no boundary term.  Since \(\delta\varphi_a^A\) is arbitrary, the
Euler--Lagrange equation is \(\partial_\mu J_A^\mu=0\).  If the microscopic
current is anomalous, the branch gauge transformation of the regulated
generating functional contains the corresponding local anomaly functional;
that case belongs to the anomalous-transport chapter, not to the neutral
diffusion action treated here.
\end{remark}

\section{Quadratic Diffusion Action}

Consider a homogeneous thermal state at temperature \(T>0\).  Let
\(\chi_{AB}\) be the static susceptibility matrix
\[
  \chi_{AB}=\left.{\frac{\partial n_A}{\partial\mu_B}}\right|_T .
\]
At a thermodynamically stable point \(\chi\) is real, symmetric, and
positive definite on the space of non-redundant conserved charges.  Let
\(\Sigma_{AB}\) be the dissipative conductivity matrix in the convention of
Chapter~\ref{ch:thermal-hydrodynamics-from-ward-identities}; for
time-reversal-even charges and in the absence of parity-odd nondissipative
transport, \(\Sigma\) is real, symmetric, and positive semidefinite.  The
diffusion endomorphism acting on densities is
\begin{equation}
  D_A{}^B=\Sigma_{AC}(\chi^{-1})^{CB}.
  \label{eq:sk-hydro-diffusion-matrix}
\end{equation}

\begin{proposition}[Stability of the multi-charge diffusion matrix]
\label{prop:sk-hydro-diffusion-matrix-stability}
Assume \(\chi\) is real symmetric positive definite and \(\Sigma\) is real
symmetric positive semidefinite.  Then
\[
  D=\Sigma\chi^{-1}
  \label{eq:sk-hydro-diffusion-endomorphism-matrix}
\]
is diagonalizable over \(\mathbb R\) and all its eigenvalues are
nonnegative.  It is self-adjoint for the density inner product
\[
  \langle u,v\rangle_{\chi^{-1}}
  =
  u_A(\chi^{-1})^{AB}v_B .
  \label{eq:sk-hydro-density-inner-product}
\]
If \(\Sigma\) is positive definite, all eigenvalues of \(D\) are positive.
\end{proposition}

\begin{proof}
Let \(\chi^{1/2}\) be the positive square root of \(\chi\).  The matrix
\[
  S=\chi^{-1/2}\Sigma\chi^{-1/2}
  \label{eq:sk-hydro-symmetric-diffusion-similarity}
\]
is real symmetric positive semidefinite.  Moreover
\[
  D=\Sigma\chi^{-1}
  =
  \chi^{1/2}S\chi^{-1/2},
  \label{eq:sk-hydro-diffusion-similar-to-symmetric}
\]
so \(D\) is similar to \(S\).  Therefore \(D\) has the same spectrum as
\(S\), is diagonalizable over \(\mathbb R\), and has nonnegative
eigenvalues.  For self-adjointness,
\[
  \langle u,Dv\rangle_{\chi^{-1}}
  =
  u^T\chi^{-1}\Sigma\chi^{-1}v
  =
  (Du)^T\chi^{-1}v
  =
  \langle Du,v\rangle_{\chi^{-1}},
  \label{eq:sk-hydro-diffusion-self-adjoint-proof}
\]
because \(\chi^{-1}\) and \(\Sigma\) are symmetric.  If \(\Sigma\) is
positive definite then \(S\) is positive definite, so its eigenvalues, and
hence the eigenvalues of \(D\), are strictly positive.
\end{proof}

The leading quadratic local action for the diffusive modes is
\begin{equation}
  I_{\rm diff}
  =
  \int dt\,d^dx\,
  \left[
  \chi_{AB}B^A_{a0}B^B_{r0}
  -\Sigma_{AB}B^A_{ai}\partial_t B^B_{ri}
  +\ii T\Sigma_{AB}B^A_{ai}B^B_{ai}
  \right],
  \label{eq:sk-hydro-diffusion-action}
\end{equation}
up to terms with more derivatives or more powers of \(a\)-fields.  The
one-charge formula is recovered by setting
\(\chi_{11}=\chi\), \(\Sigma_{11}=\sigma\), and \(D=\sigma/\chi\).  The
current obtained by varying with respect to the \(a\)-type source at
\(A^A_a=\varphi^A_a=0\) is
\begin{equation}
  n_A=\frac{\delta I_{\rm diff}}{\delta A^A_{a0}}
  =\chi_{AB}B^B_{r0},
  \qquad
  j_A^i=\frac{\delta I_{\rm diff}}{\delta A^A_{ai}}
  =-\Sigma_{AB}\partial_t B^B_{ri}.
  \label{eq:sk-hydro-density-current-from-action}
\end{equation}
Using \(B^A_{r0}=A^A_{r0}+\partial_t\varphi^A_r\) and
\(B^A_{ri}=A^A_{ri}+\partial_i\varphi^A_r\), this current may be written as
\begin{equation}
  j_A^i
  =
  -D_A{}^B\partial_i n_B+\Sigma_{AB}E^B_i,
  \qquad
  E^A_i=\partial_i A^A_{r0}-\partial_t A^A_{ri}.
  \label{eq:sk-hydro-constitutive-current-with-source}
\end{equation}
The sign of \(E^A_i\) is fixed by the source convention in which a static
scalar-source perturbation changes the density by
\(\delta n_A=\chi_{AB}\delta A^B_{r0}\).  Equation
\eqref{eq:sk-hydro-constitutive-current-with-source} is the sourceful
first-order constitutive relation
\[
  j_A^i=\Sigma_{AB}\partial_i A^B_{r0}
  -\Sigma_{AB}\partial_i\mu_B,
  \qquad
  \mu_A=B^A_{r0}=(\chi^{-1})^{AB}n_B ,
\]
written with the hydrodynamic phase made explicit.

\begin{remark}[Diffusion equation from the \(a\)-phase]
\label{rem:sk-hydro-diffusion-equation}
At \(A^A_a=\varphi^A_a=0\), varying
\eqref{eq:sk-hydro-diffusion-action} with respect to \(\varphi^A_a\) gives
\begin{equation}
  \partial_t n_A+\partial_i j_A^i=0,
  \label{eq:sk-hydro-continuity-from-phase}
\end{equation}
with \(n_A\) and \(j_A^i\) as in
\eqref{eq:sk-hydro-density-current-from-action}.  With external sources set
to zero this becomes
\begin{equation}
  \partial_t n_A-D_A{}^B\nabla^2 n_B=0 .
  \label{eq:sk-hydro-diffusion-equation}
\end{equation}

Under \(\varphi^A_a\mapsto\varphi^A_a+\delta\varphi^A_a\),
\[
  \delta B^A_{a0}=\partial_t\delta\varphi^A_a,
  \qquad
  \delta B^A_{ai}=\partial_i\delta\varphi^A_a .
\]
The terms linear in \(a\)-fields therefore vary as
\[
  \delta I_{\rm diff}
  =
  \int dt\,d^dx\,
  \left[
  \chi_{AB}(\partial_t\delta\varphi^A_a)B^B_{r0}
  -\Sigma_{AB}(\partial_i\delta\varphi^A_a)\partial_tB^B_{ri}
  \right],
\]
because the quadratic \(B_aB_a\) term does not contribute at
\(A^A_a=\varphi^A_a=0\).  After integrating by parts and discarding boundary
terms compatible with the closed-time-path variational problem,
\[
  \delta I_{\rm diff}
  =
  -\int dt\,d^dx\,\delta\varphi^A_a
  \left[
  \partial_t(\chi_{AB}B^B_{r0})
  -\partial_i(\Sigma_{AB}\partial_tB^B_{ri})
  \right].
\]
Using \eqref{eq:sk-hydro-density-current-from-action}, the Euler-Lagrange
equation is exactly \eqref{eq:sk-hydro-continuity-from-phase}.  If
\(A^A_{r\mu}=0\), then \(n_A=\chi_{AB}\partial_t\varphi^B_r\) and
\[
  j_A^i=-\Sigma_{AB}\partial_i\partial_t\varphi^B_r
  =-\Sigma_{AB}(\chi^{-1})^{BC}\partial_i n_C
  =-D_A{}^C\partial_i n_C,
\]
which gives \eqref{eq:sk-hydro-diffusion-equation}.
\end{remark}

\section{Density Response Kernel}

The density response to a scalar source follows from the same saddle
equation.  Set \(A^A_{ri}=0\) and use a Fourier mode
\(\exp(-\ii\omega t+\ii\mathbf k\cdot\mathbf x)\).  Write
\(\mathbf A_0\), \(\boldsymbol\varphi\), and \(\mathbf n\) for the column
vectors with components \(A^A_{r0}\), \(\varphi^A_r\), and \(n_A\).  The
continuity equation gives
\[
  \mathbf n+k^2\Sigma\boldsymbol\varphi=0,
  \qquad
  \mathbf n=\chi(\mathbf A_0-\ii\omega\boldsymbol\varphi).
\]
Equivalently,
\[
  (k^2\Sigma-\ii\omega\chi)\boldsymbol\varphi=-\chi\mathbf A_0 .
\]
Eliminating \(\boldsymbol\varphi\) and using \(D=\Sigma\chi^{-1}\) gives
the matrix source-response kernel
\begin{equation}
  K^R_{n_An_B}(\omega,\mathbf k)
  =
  \frac{\delta n_A}{\delta A^B_{r0}}
  =
  \left[
    (Dk^2-\ii\omega\mathbf 1)^{-1}k^2\Sigma
  \right]_{AB}.
  \label{eq:sk-hydro-density-response-kernel}
\end{equation}
It has the diffusive pole
\begin{equation}
  \det(Dk^2-\ii\omega\mathbf 1)=0 .
  \label{eq:sk-hydro-density-response-pole}
\end{equation}
At fixed nonzero \(k\), the static limit is
\[
  K^R_{n_An_B}(0,\mathbf k)=\chi_{AB},
\]
provided \(D\) is invertible on the charge sector being probed.  If
\(\Sigma\) has an exact null vector, the corresponding charge does not
diffuse at this derivative order and must be treated by retaining the next
allowed term.  At \(k=0\) and nonzero \(\omega\), the kernel vanishes,
expressing conservation of the total charges.  These two limits are
distinct; the hydrodynamic poles interpolate between them.

For a transverse vector source, \(k_iA^A_{ri}=0\) and \(A^A_{r0}=0\), the
phase decouples at quadratic order and
\[
  j_A^i=-\Sigma_{AB}\partial_tA^B_{ri}.
\]
With the electric-field convention
\eqref{eq:sk-hydro-constitutive-current-with-source}, this is Ohm response
\(j_A^i=\Sigma_{AB}E^B_i\).  The same matrix \(\Sigma\) therefore appears in
the diffusion endomorphism, the vector-source response, and the noise
covariance below.

\section{Noise Positivity and Langevin Form}

The imaginary part of \eqref{eq:sk-hydro-diffusion-action} is
\begin{equation}
  \operatorname{Im} I_{\rm diff}
  =
  T\int dt\,d^dx\,\Sigma_{AB}B^A_{ai}B^B_{ai}.
  \label{eq:sk-hydro-imaginary-action}
\end{equation}
It is nonnegative precisely when \(\Sigma\) is positive semidefinite.  The
same term has a stochastic representation on the image of \(\Sigma\).  If
\(\Sigma\) is positive definite, then for each spatial component
\begin{equation}
  \exp\!\left[
    -T\int\Sigma_{AB}B^A_{ai}B^B_{ai}
  \right]
  =
  \int D\xi\,
  \exp\!\left[
    -\int\frac{1}{4T}
      \xi_{Ai}(\Sigma^{-1})^{AB}\xi_{Bi}
    +\ii\int B^A_{ai}\xi_{Ai}
  \right],
  \label{eq:sk-hydro-hubbard-stratonovich}
\end{equation}
with normalization independent of \(B_a\).  For positive semidefinite
\(\Sigma\), the same formula is understood after restricting to the support
of \(\Sigma\), or as a limit from positive definite matrices.  Thus the
noise satisfies
\begin{equation}
  \langle \xi_{Ai}(t,\mathbf x)\xi_{Bj}(t',\mathbf x')\rangle
  =
  2T\Sigma_{AB}\,\delta_{ij}
  \delta(t-t')\delta^{(d)}(\mathbf x-\mathbf x').
  \label{eq:sk-hydro-noise-correlator}
\end{equation}
After the Hubbard--Stratonovich transformation, varying \(\varphi^A_a\) gives
the Langevin equation
\begin{equation}
  \partial_t n_A+
  \partial_i\bigl(-D_A{}^B\partial_i n_B+\Sigma_{AB}E^B_i+\xi_{Ai}\bigr)=0 .
  \label{eq:sk-hydro-langevin-diffusion}
\end{equation}
This derivation fixes the factor \(2T\Sigma_{AB}\) in the noise correlator
from the same matrix that governs the dissipative current.

\begin{theorem}[Density fluctuation--dissipation theorem from the diffusion action]
\label{thm:sk-hydro-density-fdt-from-action}
Assume \(\chi\) is positive definite and \(\Sigma\) is positive semidefinite.
For real \(\omega\), \(\mathbf k\), let
\[
  M(\omega,\mathbf k)=
  (Dk^2-\ii\omega\mathbf 1)^{-1},
  \qquad k^2=\mathbf k\cdot\mathbf k ,
  \label{eq:sk-hydro-diffusion-resolvent}
\]
on the subspace where the inverse exists.  The Langevin equation
\eqref{eq:sk-hydro-langevin-diffusion} at zero external source gives the
symmetrized density correlator
\[
  G^{S}_{nn}(\omega,\mathbf k)
  =
  2T\,k^2\,
  M(\omega,\mathbf k)\,
  \Sigma\,
  M(-\omega,\mathbf k)^T .
  \label{eq:sk-hydro-density-symmetrized-correlator}
\]
This matrix is positive semidefinite as a Hermitian kernel.  If
\[
  K^A_{nn}(\omega,\mathbf k)
  =
  K^R_{nn}(-\omega,\mathbf k)^T
  =
  k^2\Sigma\,M(-\omega,\mathbf k)^T ,
  \label{eq:sk-hydro-density-advanced-kernel}
\]
then, for \(\omega\ne0\),
\[
  G^{S}_{nn}(\omega,\mathbf k)
  =
  \frac{2T}{\omega}\,
  \frac{K^R_{nn}(\omega,\mathbf k)-K^A_{nn}(\omega,\mathbf k)}{2\ii}.
  \label{eq:sk-hydro-density-classical-fdt}
\]
The equality is the classical low-frequency form of the KMS
fluctuation--dissipation relation in the density sector.
\end{theorem}

\begin{proof}
Set the external source to zero.  With the Fourier convention fixed at the
beginning of the chapter,
\eqref{eq:sk-hydro-langevin-diffusion} becomes
\[
  (Dk^2-\ii\omega\mathbf 1)\,\mathbf n
  =
  -\ii k_i\,\boldsymbol\xi_i .
  \label{eq:sk-hydro-langevin-fourier-saddle}
\]
Thus \(\mathbf n=-\ii k_iM\boldsymbol\xi_i\).  Using
\eqref{eq:sk-hydro-noise-correlator} and summing over the spatial noise
index gives exactly
\eqref{eq:sk-hydro-density-symmetrized-correlator}.  For real \(\omega\) and
real \(D,\Sigma\), \(M(-\omega)^T=M(\omega)^\dagger\) after lowering and
raising only ordinary finite charge indices, so
\[
  u^\dagger G^S_{nn}u
  =
  2Tk^2\,(M^\dagger u)^\dagger\Sigma(M^\dagger u)\ge0
  \label{eq:sk-hydro-density-correlator-positivity}
\]
for every complex vector \(u\).

It remains to prove the matrix fluctuation--dissipation identity without
assuming that \(D\) and \(\Sigma\) commute.  The response kernel from
\eqref{eq:sk-hydro-density-response-kernel} is
\[
  K^R_{nn}=M k^2\Sigma .
  \label{eq:sk-hydro-density-retarded-matrix-proof}
\]
Since \(D=\Sigma\chi^{-1}\) and \(\chi,\Sigma\) are symmetric,
\[
  (Dk^2-\ii\omega)\chi=k^2\Sigma-\ii\omega\chi,
  \qquad
  \chi(D^Tk^2+\ii\omega)=k^2\Sigma+\ii\omega\chi .
  \label{eq:sk-hydro-fdt-chi-intertwining}
\]
Therefore
\[
  K^R_{nn}
  =
  M\bigl[(Dk^2-\ii\omega)\chi+\ii\omega\chi\bigr]
  =
  \chi+\ii\omega M\chi ,
  \label{eq:sk-hydro-fdt-retarded-rewrite}
\]
and similarly
\[
  K^A_{nn}
  =
  \bigl[\chi(D^Tk^2+\ii\omega)-\ii\omega\chi\bigr]M(-\omega)^T
  =
  \chi-\ii\omega\chi M(-\omega)^T .
  \label{eq:sk-hydro-fdt-advanced-rewrite}
\]
Subtracting gives
\[
  K^R_{nn}-K^A_{nn}
  =
  \ii\omega\bigl(M\chi+\chi M(-\omega)^T\bigr).
  \label{eq:sk-hydro-fdt-difference-first}
\]
Multiplying
\eqref{eq:sk-hydro-fdt-chi-intertwining} on the left and right by the two
resolvents gives
\[
  M\chi+\chi M(-\omega)^T
  =
  2k^2M\Sigma M(-\omega)^T .
  \label{eq:sk-hydro-fdt-resolvent-identity}
\]
Combining the last two displays yields
\[
  \frac{K^R_{nn}-K^A_{nn}}{2\ii}
  =
  \omega k^2M\Sigma M(-\omega)^T ,
  \label{eq:sk-hydro-fdt-spectral-density}
\]
which is equivalent to
\eqref{eq:sk-hydro-density-classical-fdt} after using
\eqref{eq:sk-hydro-density-symmetrized-correlator}.
\end{proof}

\section{Dynamical KMS at Quadratic Order}

Assume the initial state is thermal with inverse temperature
\(\beta=1/T\) and that the long-wavelength effective action has the
classical dynamical KMS symmetry appropriate to the derivative order under
consideration.  In the diffusion sector, and for time-reversal-even charge
densities, the transformation needed at this order is
\begin{equation}
  B^A_{ai}(t,\mathbf x)
  \mapsto
  B^A_{ai}(-t,\mathbf x)
  +\ii\beta\,\partial_tB^A_{ri}(-t,\mathbf x),
  \label{eq:sk-hydro-classical-kms-ba}
\end{equation}
with the time-reversal parity of the spatial vector field included in
\(\partial_tB^A_{ri}(-t,\mathbf x)\).  To see the coefficient constraint,
keep one spatial component and write the dissipative-noise part as
\[
  L=-a^A\Sigma_{AB}x^B+\ii a^AC_{AB}a^B,
  \qquad
  a^A=B^A_{ai},
  \qquad
  x^A=\partial_tB^A_{ri},
\]
where \(C_{AB}\) is the unknown symmetric noise matrix.  Time reversal sends
\(x\mapsto -x\), while the KMS shift sends
\(a\mapsto a+\ii\beta x\).  The transformed expression is
\[
  L'
  =
  a^A(\Sigma_{AB}-2\beta C_{AB})x^B
  +\ii a^AC_{AB}a^B
  +\ii\beta x^A(\Sigma_{AB}-\beta C_{AB})x^B .
\]
Equality with \(L\), up to total derivatives and terms beyond the retained
order, requires the coefficient of \(a^Ax^B\) to match:
\[
  \Sigma_{AB}-2\beta C_{AB}=-\Sigma_{AB}.
\]
The \(x^Ax^B\) terms then cancel by the same condition.  Therefore
\begin{equation}
  C_{AB}=\frac{\Sigma_{AB}}{\beta}=T\Sigma_{AB},
  \label{eq:sk-hydro-kms-noise-coefficient}
\end{equation}
which is the coefficient in \eqref{eq:sk-hydro-diffusion-action}.  This is
the local effective-action form of the fluctuation--dissipation relation of
Chapter~\ref{ch:thermal-spectral-kubo-transport}.

\section{Relation to Stress-Tensor Hydrodynamics}

For stress-tensor hydrodynamics, the charge phase fields are replaced by
doubled maps from a fluid worldvolume into spacetime, together with doubled
metrics and, when charges are present, the phases above.  The structural
logic is the same:
\[
\begin{aligned}
  \text{gauge/diffeomorphism invariance}
  &\Longrightarrow
  \text{Ward identities},\\[2mm]
  I[A,A]=0
  &\Longrightarrow
  \text{closed-time-path normalization},\\[2mm]
  \operatorname{Im}I\ge0
  &\Longrightarrow
  \text{positive noise kernel},\\[2mm]
  \text{dynamical KMS}
  &\Longrightarrow
  \text{dissipation--fluctuation relations}.
\end{aligned}
\]
At first derivative order this reproduces the viscous stress tensor and
noise kernel used in Chapter~\ref{ch:thermal-hydrodynamic-fluctuations}.
At higher derivative order the classification becomes a local invariant
problem for doubled fields subject to the constraints above.  A term belongs
to the hydrodynamic effective action only after its frame dependence,
contact-term ambiguity, KMS transformation, and positivity properties have
been specified.

\section{The Microscopic Theorem Boundary}

The construction in this chapter is an effective-field-theory statement.  A
microscopic derivation from QFT would have to construct the real-time
generating functional in a thermal state, take the thermodynamic limit,
identify the hydrodynamic scaling limit, prove that the only nonanalytic
long-wavelength modes are the conserved densities, and show that the local
functional above is the asymptotic expansion of connected
Schwinger--Keldysh correlators in a controlled topology.  The local
positivity and KMS constraints are necessary consequences of the microscopic
closed time path; they are not by themselves a proof of thermalization.

\begin{openproblem}[Microscopic derivation of SK hydrodynamic action]
\label{op:microscopic-sk-hydrodynamic-action}
Starting from a continuum QFT with a KMS state, derive the hydrodynamic
Schwinger--Keldysh effective action as a controlled long-wavelength limit of
connected real-time correlators, including the topology in which the
derivative expansion is asymptotic, the treatment of contact terms, and the
relation between source-response kernels and commutator spectral functions.
\end{openproblem}
