\chapter{Real-Time Schwinger--Keldysh Formalism}
\label{ch:thermal-schwinger-keldysh}

\ConventionsRealTime

The Schwinger--Keldysh formalism is the real-time generating functional for
correlation and response functions in a specified initial state.  It is an
operator identity before it is a path integral.  The doubled fields and
sources are a bookkeeping device for the two sides of a trace, and the
fundamental constraints come from unitarity, positivity of the density
operator, and, in thermal equilibrium, the KMS condition.

\section{Finite-System Definition}

Let \(\Hilb\) be a Hilbert space, let \(H\) be a self-adjoint Hamiltonian, and
let \(\rho_0\) be a positive trace-class density operator with
\[
  \operatorname{Tr}\rho_0=1 .
\]
Let \(O_\alpha\), \(\alpha\in\mathcal I\), be a finite or regulated family of
Hermitian operators.  For real compactly supported sources \(J^\alpha(t)\),
define the source-dependent Hamiltonian
\[
  H_J(t)=H-\sum_{\alpha\in\mathcal I} J^\alpha(t)O_\alpha .
\]
The evolution operator \(U_J(t,t_i)\) is the solution of
\[
  \ii\frac{d}{dt}U_J(t,t_i)=H_J(t)U_J(t,t_i),
  \qquad U_J(t_i,t_i)=\mathbf 1 ,
\]
equivalently
\[
  U_J(t_f,t_i)
  =
  \operatorname{T}\exp\!\left[
  -\ii\int_{t_i}^{t_f}dt\,
  \left(H-\sum_{\alpha}J^\alpha(t)O_\alpha\right)\right].
\]
The closed-time-path generating functional is
\begin{equation}
  Z[J_+,J_-]
  =
  \operatorname{Tr}\!\left(
    U_{J_+}(t_f,t_i)\rho_0
    U_{J_-}(t_f,t_i)^\dagger
  \right).
  \label{eq:sk-operator-generating-functional}
\end{equation}
The \(+\) and \(-\) labels record whether a source acts on the forward or
backward side of the trace.  The physical source is obtained on the diagonal
\(J_+=J_-\).

\begin{definition}[Closed-time-path correlators]
\label{def:closed-time-path-correlators}
The Schwinger--Keldysh correlators of the regulated system are the
functional derivatives of \(Z[J_+,J_-]\) at \(J_+=J_-=0\).  Differentiation
with respect to \(J_+\) inserts operators on the forward branch with time
ordering; differentiation with respect to \(J_-\) inserts operators on the
backward branch with anti-time ordering and an opposite sign from the
adjoint evolution.
\end{definition}

For example, in the unperturbed Heisenberg convention
\[
  O_\alpha(t)=\ee^{\ii H(t-t_i)}O_\alpha\ee^{-\ii H(t-t_i)},
\]
one has
\[
  \left.
  \frac{\delta Z}{\delta J_+^\alpha(t)}
  \right|_{0}
  =
  \ii\,\operatorname{Tr}(\rho_0 O_\alpha(t)),
  \qquad
  \left.
  \frac{\delta Z}{\delta J_-^\alpha(t)}
  \right|_{0}
  =
  -\ii\,\operatorname{Tr}(\rho_0 O_\alpha(t)).
\]
Higher derivatives give the four contour orderings.  With two insertions,
\begin{align}
  \left.
  \frac{1}{\ii^2}
  \frac{\delta^2Z}{\delta J_+^\alpha(t)\delta J_+^\beta(t')}
  \right|_{0}
  &=
  \operatorname{Tr}\!\left(
    \rho_0\,\operatorname{T}O_\alpha(t)O_\beta(t')\right),\\
  \left.
  \frac{1}{(-\ii)^2}
  \frac{\delta^2Z}{\delta J_-^\alpha(t)\delta J_-^\beta(t')}
  \right|_{0}
  &=
  \operatorname{Tr}\!\left(
    \rho_0\,\widetilde{\operatorname{T}}O_\alpha(t)O_\beta(t')\right),\\
  \left.
  \frac{1}{\ii(-\ii)}
  \frac{\delta^2Z}{\delta J_+^\alpha(t)\delta J_-^\beta(t')}
  \right|_{0}
  &=
  \operatorname{Tr}\!\left(\rho_0\,O_\beta(t')O_\alpha(t)\right),\\
  \left.
  \frac{1}{(-\ii)\ii}
  \frac{\delta^2Z}{\delta J_-^\alpha(t)\delta J_+^\beta(t')}
  \right|_{0}
  &=
  \operatorname{Tr}\!\left(\rho_0\,O_\alpha(t)O_\beta(t')\right).
\end{align}
These formulae are finite trace identities.  In a continuum QFT the same
symbols require a regulator, an operator topology, and a limiting procedure.

\section{Unitarity, Reality, and Positivity Constraints}

Setting the two sources equal gives
\begin{equation}
  Z[J,J]
  =
  \operatorname{Tr}(U_J\rho_0U_J^\dagger)
  =
  \operatorname{Tr}\rho_0
  =
  1 .
  \label{eq:sk-unitarity-normalization}
\end{equation}
This is the closed-time-path normalization constraint.

\begin{proposition}[Basic Schwinger--Keldysh constraints]
\label{prop:basic-sk-constraints}
For real sources in the finite system above,
\begin{align}
  Z[J,J]&=1,\\
  Z[J_+,J_-]^*&=Z[J_-,J_+],\\
  |Z[J_+,J_-]|&\leq1 .
\end{align}
If \(Z=\exp(\ii W)\) is defined by the branch with \(W[0,0]=0\), then near
the origin
\[
  W[J,J]=0,\qquad
  W[J_+,J_-]^*=-W[J_-,J_+],\qquad
  \operatorname{Im}W[J_+,J_-]\geq0 .
\]
\end{proposition}

\begin{proof}
The first identity is \eqref{eq:sk-unitarity-normalization}.  For real
sources \(U_J\) is unitary, and complex conjugation of the trace gives
\[
  Z[J_+,J_-]^*
  =
  \operatorname{Tr}(U_{J_-}\rho_0U_{J_+}^\dagger)
  =
  Z[J_-,J_+] .
\]
For the bound, write
\[
  Z[J_+,J_-]
  =
  \operatorname{Tr}\!\left(
    \rho_0\,U_{J_-}^\dagger U_{J_+}
  \right).
\]
The operator \(V=U_{J_-}^\dagger U_{J_+}\) is unitary.  Since \(\rho_0\) is a
positive normalized density matrix, \(Z=\omega_{\rho_0}(V)\) is an
expectation value of a unitary in a state.  Cauchy--Schwarz for the positive
inner product
\[
  (A,B)_{\rho_0}=\operatorname{Tr}(\rho_0 A^\dagger B)
\]
gives
\[
  |Z|^2\leq
  \operatorname{Tr}(\rho_0\mathbf 1)
  \operatorname{Tr}(\rho_0 V^\dagger V)=1.
\]
The statements about \(W\) follow from \(Z=\exp(\ii W)\) on the branch
connected to \(Z=1\).
\end{proof}

Equation \eqref{eq:sk-unitarity-normalization} implies that all derivatives
of \(Z\) or \(W\) with respect only to a common physical source vanish on the
diagonal.  This is the origin of the rule that every local term in a
Schwinger--Keldysh effective action contains at least one difference field.

\section{The r/a Basis and Retarded Response}

Introduce average and difference sources
\[
  J_{\mathrm r}^\alpha=\frac12(J_+^\alpha+J_-^\alpha),
  \qquad
  J_{\mathrm a}^\alpha=J_+^\alpha-J_-^\alpha .
\]
Equivalently,
\[
  J_+^\alpha=J_{\mathrm r}^\alpha+\frac12J_{\mathrm a}^\alpha,
  \qquad
  J_-^\alpha=J_{\mathrm r}^\alpha-\frac12J_{\mathrm a}^\alpha .
\]
The connected generating functional is
\[
  W[J_{\mathrm r},J_{\mathrm a}]
  =-\ii\log Z[J_{\mathrm r},J_{\mathrm a}],
  \qquad W[0,0]=0 .
\]
The expectation value in the physical source \(J_{\mathrm r}\) is
\begin{equation}
  \langle O_\alpha(t)\rangle_{J_{\mathrm r}}
  =
  \left.
  \frac{\delta W}{\delta J_{\mathrm a}^\alpha(t)}
  \right|_{J_{\mathrm a}=0}.
  \label{eq:sk-one-point-from-a-source}
\end{equation}
At zero source this follows from the one-point identities above:
\[
  \left.
  \frac{\delta Z}{\delta J_{\mathrm a}^\alpha(t)}
  \right|_0
  =
  \frac12\left(
  \ii\langle O_\alpha(t)\rangle+\ii\langle O_\alpha(t)\rangle
  \right)
  =
  \ii\langle O_\alpha(t)\rangle .
\]

\begin{proposition}[Source-response kernel from \(r/a\) differentiation]
\label{prop:sk-retarded-kernel}
With the source convention \(H_J=H-J^\alpha O_\alpha\),
\[
  \left.
  \frac{\delta}{\delta J_{\mathrm r}^\beta(t')}
  \frac{\delta W}{\delta J_{\mathrm a}^\alpha(t)}
  \right|_{J_{\mathrm r}=J_{\mathrm a}=0}
  =
  K^R_{\alpha\beta}(t,t'),
\]
where
\[
  K^R_{\alpha\beta}(t,t')
  =
  \ii\,\theta(t-t')\,
  \operatorname{Tr}\!\left(\rho_0[O_\alpha(t),O_\beta(t')]\right).
\]
If the commutator-retarded convention is
\[
  G^{R,\mathrm{comm}}_{\alpha\beta}(t,t')
  =
  -\ii\,\theta(t-t')\,
  \operatorname{Tr}\!\left(\rho_0[O_\alpha(t),O_\beta(t')]\right),
\]
then \(K^R_{\alpha\beta}=-G^{R,\mathrm{comm}}_{\alpha\beta}\).  The sign is
fixed by the source coupling \(H_J=H-J^\gamma O_\gamma\), with summation over
\(\gamma\in\mathcal I\).
\end{proposition}

\begin{proof}
Let \(J_{\mathrm r}\) be the same physical source on both branches.  To first
order in the source, the interaction-picture evolution gives
\[
  U_J(t,t_i)
  =
  U_0(t,t_i)
  \left[
    1+\ii\int_{t_i}^{t}ds\,J_{\mathrm r}^\beta(s)O_\beta(s)
    +O(J_{\mathrm r}^2)
  \right].
\]
Inserting this in the Heisenberg expectation of \(O_\alpha(t)\) gives
\[
\begin{aligned}
  \delta\langle O_\alpha(t)\rangle
  &=
  \ii\int_{t_i}^{t}ds\,J_{\mathrm r}^\beta(s)\,
  \operatorname{Tr}\!\left(\rho_0[O_\alpha(t),O_\beta(s)]\right)\\
  &=
  \int ds\,J_{\mathrm r}^\beta(s)\,
  \bigl[\ii\theta(t-s)
  \operatorname{Tr}(\rho_0[O_\alpha(t),O_\beta(s)])\bigr].
\end{aligned}
\]
Combining this linear-response identity with
\eqref{eq:sk-one-point-from-a-source} gives the displayed derivative of
\(W\).
\end{proof}

The same statement can be read directly from contour ordering.  The
derivative with respect to \(J_{\mathrm a}\) inserts the physical operator
whose expectation value is measured; the derivative with respect to
\(J_{\mathrm r}\) inserts the perturbation.  The cancellation between the two
contour branches removes all terms in which the perturbation is later than
the measured operator, leaving the causal source-response commutator
\(\ii\theta(t-t')\operatorname{Tr}(\rho_0[O_\alpha(t),O_\beta(t')])\).

\section{Two-Point Matrix}

For a bosonic Hermitian operator \(O\), define
\[
  G^>(t,t')=\operatorname{Tr}(\rho_0O(t)O(t')),
  \qquad
  G^<(t,t')=\operatorname{Tr}(\rho_0O(t')O(t)).
\]
The contour-ordered functions are
\begin{align}
  G_{++}(t,t')
  &=
  \theta(t-t')G^>(t,t')+\theta(t'-t)G^<(t,t'),\\
  G_{--}(t,t')
  &=
  \theta(t'-t)G^>(t,t')+\theta(t-t')G^<(t,t'),\\
  G_{-+}(t,t')&=G^>(t,t'),\\
  G_{+-}(t,t')&=G^<(t,t').
\end{align}
In the \(r/a\) basis the independent linear-response and fluctuation data are
\[
  G^{R,\mathrm{comm}}(t,t')=
  -\ii\theta(t-t')\bigl(G^>(t,t')-G^<(t,t')\bigr),
\]
\[
  G^{A,\mathrm{comm}}(t,t')=
  \ii\theta(t'-t)\bigl(G^>(t,t')-G^<(t,t')\bigr),
\]
\[
  G^K(t,t')=G^>(t,t')+G^<(t,t').
\]
The component with two difference-type operator insertions vanishes:
\[
  G^{\mathrm{aa}}(t,t')=0 .
\]
This is another form of
\(Z[J_{\mathrm r},J_{\mathrm a}=0]=1\).  The \(rr\) component is the
symmetrized fluctuation, while \(ra\) and \(ar\) contain the causal response
kernels.  With the source convention \(H-J^\alpha O_\alpha\), the physical
source-response kernel is the negative of \(G^{R,\mathrm{comm}}\).

\section{Thermal KMS Constraint on the Contour}

Let
\[
  \rho_\beta=Z(\beta)^{-1}\ee^{-\beta H}.
\]
For operators whose thermal strip analytic continuation is controlled,
cyclicity of the trace gives
\[
  \operatorname{Tr}(\rho_\beta A(t)B(0))
  =
  \operatorname{Tr}(\rho_\beta B(0)A(t+\ii\beta)).
\]
In frequency space this becomes detailed balance:
\[
  G^<(\omega)=\ee^{-\beta\omega}G^>(\omega)
\]
for bosonic operators in the convention of
Chapter~\ref{ch:kms-states-thermal-correlators}.  Therefore
\[
  G^K(\omega)
  =
  \coth\!\left(\frac{\beta\omega}{2}\right)
  \bigl(G^>(\omega)-G^<(\omega)\bigr).
\]
Since
\[
  G^{R,\mathrm{comm}}(\omega)-G^{A,\mathrm{comm}}(\omega)
  =
  -\ii\bigl(G^>(\omega)-G^<(\omega)\bigr)
\]
with the commutator-retarded convention above, the fluctuation-dissipation
relation is
\[
  G^K(\omega)
  =
  \ii\coth\!\left(\frac{\beta\omega}{2}\right)
  \bigl(G^{R,\mathrm{comm}}(\omega)
  -G^{A,\mathrm{comm}}(\omega)\bigr).
\]
For fermionic operators the graded KMS condition replaces the hyperbolic
cotangent by the hyperbolic tangent in the corresponding anticommutator
fluctuation function.

Thermal Schwinger--Keldysh path integrals often implement this condition by
adding an imaginary Euclidean segment of length \(\beta\) to the contour or,
in effective theory, by imposing a dynamical KMS symmetry.  Both
formulations express the same operator statement: the initial density matrix
is the Gibbs operator and the required strip analyticity exists.

\section{Coordinate Path Integrals}

A Schwinger--Keldysh path integral is a coordinate representation of
\eqref{eq:sk-operator-generating-functional}.  For a regulated bosonic
configuration variable \(q\), inserting coordinate resolutions of identity
on both sides of the trace gives two histories
\[
  q_+(t),\qquad q_-(t),
\]
with a final-time gluing condition implementing the trace:
\[
  q_+(t_f)=q_-(t_f).
\]
The initial density matrix contributes its coordinate kernel
\[
  \rho_0(q_+(t_i),q_-(t_i)).
\]
In formal continuum notation,
\[
  Z[J_+,J_-]
  =
  \int_{\mathrm{SK}}Dq_+Dq_-\,
  \rho_0(q_+(t_i),q_-(t_i))\,
  \exp\!\left\{
  \ii S[q_+]-\ii S[q_-]
  +\ii\int J_+O[q_+]-\ii\int J_-O[q_-]
  \right\}.
\]
This formula has the same logical status as every path integral in the
monograph: its meaning is supplied by the finite regulator or constructive
definition from which it is obtained.  For thermal initial data, the kernel
\(\rho_\beta\) may be represented by the Euclidean thermal path integral of
Chapter~\ref{ch:finite-temperature-path-integrals}, producing the familiar
contour with an imaginary thermal segment.

\section{Gauge Theories and BV Data on the Closed Time Path}

For a gauge theory, branch doubling must be applied to the full gauge-fixed
or BV field space.  A perturbative BV representation has two copies
\[
  \Phi_+,\Phi_+^*,
  \qquad
  \Phi_-,\Phi_-^*,
\]
and a regulated exponent whose local part has the form
\[
  \ii S_{\rm BV}[\Phi_+,\Phi_+^*]
  -
  \ii S_{\rm BV}[\Phi_-,\Phi_-^*]
  +
  \ii J_+\cdot O[\Phi_+]
  -
  \ii J_-\cdot O[\Phi_-]
\]
plus the initial density matrix and the final gluing of physical variables.
Gauge consistency is the statement that the doubled measure, action, sources,
and boundary conditions satisfy the appropriate quantum BV master equation
and preserve the diagonal identity \(Z[J,J]=1\).  In a lattice gauge
regulator the same physical constraints are expressed without gauge fixing by
two copies of link variables on the closed time path, projection to
gauge-invariant states or Gauss-law sectors, and final gluing inside the
trace.

The real-time continuum construction is therefore stronger than writing a
formal doubled action.  It must define the initial state, operator domains,
source space, contour gluing, gauge constraints, and limiting topology in
which unitarity, positivity, KMS relations, and retarded support survive.

\begin{openproblem}[Real-time continuum construction]
\label{op:real-time-sk-continuum-construction}
Construct Schwinger--Keldysh generating functionals for interacting continuum
QFTs in a topology strong enough to derive retarded correlators, KMS
relations, and transport limits.  The construction must state the initial
state, operator domains, regulator, contour matching, gauge or BV data when
present, and the sense in which unitarity and positivity survive the
continuum limit.
\end{openproblem}
