\chapter{Anomalous and Topological Transport}
\label{ch:thermal-anomalous-topological-transport}

\ConventionsMixed

This chapter studies transport terms fixed by anomalies and topology.  The
logic is real-time response first: a transport coefficient is defined by a
correlator or by the response of a generating functional to background
sources, and anomaly equations constrain the allowed terms.

\section{Hydrodynamic Source Datum}

Let \(W[g,A]\) be the equilibrium generating functional for a QFT with
stress tensor \(T^{\mu\nu}\) and current \(J^\mu\).  Variations define
\[
  \delta W
  =
  \frac12\int\dd^Dx\,\sqrt{-g}\,
  \langle T^{\mu\nu}\rangle\,\delta g_{\mu\nu}
  +
  \int\dd^Dx\,\sqrt{-g}\,
  \langle J^\mu\rangle\,\delta A_\mu .
\]
Hydrodynamic variables are a temperature \(T\), velocity \(u^\mu\), and
chemical potential \(\mu\), with
\[
  u^\mu u_\mu=-1.
\]
The ideal constitutive relations are
\[
  T^{\mu\nu}_{(0)}=\epsilon u^\mu u^\nu+p\Delta^{\mu\nu},\qquad
  J^\mu_{(0)}=n u^\mu,
\]
where
\[
  \Delta^{\mu\nu}=g^{\mu\nu}+u^\mu u^\nu .
\]

\section{Anomaly Equation}

In four dimensions, a \(U(1)^3\) anomaly has the local equation
\[
  \nabla_\mu J^\mu
  =
  \frac{C}{32\pi^2}
  \epsilon^{\mu\nu\rho\sigma}F_{\mu\nu}F_{\rho\sigma}.
\]
The coefficient \(C\) is part of the current normalization.  The magnetic
field and vorticity in the local rest frame are
\[
  B^\mu=\frac12\epsilon^{\mu\nu\rho\sigma}u_\nu F_{\rho\sigma},
  \qquad
  \omega^\mu=\frac12\epsilon^{\mu\nu\rho\sigma}
  u_\nu\nabla_\rho u_\sigma .
\]
The first-derivative parity-odd current terms have the form
\[
  J^\mu_{\mathrm{odd}}
  =
  \xi_B B^\mu+\xi_\omega\omega^\mu .
\]
The coefficients \(\xi_B,\xi_\omega\) are transport coefficients.  They must
be defined either by Kubo formulae or by differentiating an equilibrium
generating functional.

For the explicit four-dimensional formulas below, let \(J_R\) and \(J_L\) be
the source-normalized right- and left-handed number currents of one massless
Dirac fermion, and set
\[
  J_V=J_R+J_L,\qquad J_A=J_R-J_L,\qquad
  \mu_R=\mu_V+\mu_A,\qquad \mu_L=\mu_V-\mu_A .
\]
When these Weyl sums describe a populated state, \(\mu_A\) is the
occupation-imbalance parameter.  A background axial thermal holonomy is
written separately as \(a_0^5=(A_A)_0\).
The vector source is \(A_V\).  In the convention of
Chapter~\ref{ch:volume-ii-20}, preserving the vector Ward identity gives
\[
  \nabla_\mu J_A^\mu
  =
  \frac1{16\pi^2}
  \epsilon^{\mu\nu\rho\sigma}
  (F_V)_{\mu\nu}(F_V)_{\rho\sigma}
  \qquad
  \hbox{for a unit source charge.}
\]
If a physical electromagnetic source \(A_{\rm em}\) couples with charge
\(Q=eq\), then \(A_V=Q A_{\rm em}\).  The current conjugate to
\(A_{\rm em}\) is \(QJ_V\), and the magnetic response acquires the factor
\(Q^2=e^2q^2\).

\section{Anomalous Source Protocols and Current Representatives}

Anomaly-induced transport is meaningful only after the source protocol and
the current representative have been specified.  This point is not a matter
of taste: anomalous Ward identities, Chern--Simons contact terms, and
zero-frequency limits mix under local counterterms.

\begin{definition}[Hydrostatic anomalous-source datum]
\label{def:hydrostatic-anomalous-source-datum}
A hydrostatic anomalous-source datum consists of the following objects:
\begin{enumerate}
  \item a stationary Lorentzian background \((g,A_I)\) with a timelike
  Killing field \(K\);
  \item chemical potentials \(\mu_I\) defined as the gauge-invariant thermal
  holonomies of \(A_I\) around the Euclidean thermal circle, after choosing
  the time-independent gauge used in the hydrostatic functional;
  \item a choice of local counterterms, equivalently a choice of consistent
  current representatives;
  \item a declaration of which gauge symmetries are exact constraints on the
  background functional and which sources are allowed to have anomalous Ward
  identities.
\end{enumerate}
For an anomalous charge, \(\mu_I\) is a background source parameter in this
datum.  It is a genuine grand-canonical chemical potential for a conserved
operator only in source sectors for which the corresponding anomaly density
vanishes, or after passing to a conserved combination of charges.
\end{definition}

\begin{definition}[Consistent and covariant currents]
\label{def:consistent-covariant-currents-transport}
Let \(W[A]\) be a renormalized connected generating functional.  The
consistent current is the variational current
\[
  J_{\rm cons}^\mu(x)
  =
  \frac{1}{\sqrt{-g(x)}}\frac{\delta W}{\delta A_\mu(x)}.
\]
With the sign convention
\begin{equation}
  \delta_\alpha W
  =
  -\int\dd^4x\,\sqrt{-g}\,\alpha\,\mathcal A_{\rm cons},
  \qquad
  \delta_\alpha A_\mu=\nabla_\mu\alpha ,
  \label{eq:transport-consistent-anomaly-sign}
\end{equation}
integration by parts gives
\[
  \nabla_\mu J_{\rm cons}^\mu=\mathcal A_{\rm cons}.
\]
A covariant current is a shifted local current
\begin{equation}
  J_{\rm cov}^\mu
  =
  J_{\rm cons}^\mu+J_{\rm BZ}^\mu,
  \label{eq:transport-bz-current-shift}
\end{equation}
where \(J_{\rm BZ}^\mu\) is a local Bardeen--Zumino polynomial in the
background fields chosen so that \(J_{\rm cov}\) transforms covariantly under
the background gauge symmetry being represented.  The shift
\eqref{eq:transport-bz-current-shift} changes local contact terms in current
correlators; it is not a change of the Hilbert-space operator algebra.
\end{definition}

\begin{remark}[Vector-preserving convention in the Dirac example]
\label{rem:transport-vector-preserving-bardeen}
The Dirac formulas in this chapter use the Bardeen convention in which the
vector source \(A_V\) is an exact background gauge field and the axial source
may have an anomalous Ward identity.  Thus the physical electromagnetic
response is obtained by varying the vector-preserving functional with respect
to \(A_{\rm em}\).  In the chiral-magnetic formulas below the symbol
\(a_0^5=(A_A)_0\) denotes an axial background source, while
\(\mu_5^{\rm occ}=(\mu_R-\mu_L)/2\) denotes an occupation imbalance in the
density matrix.  A conserved axial charge in a sector with nonzero
\(\epsilon F_VF_V\) requires additional anomaly-free data.
\end{remark}

\section{Kubo Definitions}

In flat space at temperature \(T\), define retarded correlators
\[
  G^{R}_{AB}(t,\mathbf x)
  =
  -\ii\theta(t)\langle[A(t,\mathbf x),B(0)]\rangle_\beta .
\]
For the vector source we use the Lorentzian source convention
\[
  \delta S_{\rm L}
  =
  \int\dd^4x\,J_V^\mu\,\delta(A_V)_\mu ,
  \qquad
  H[A_V]=H[0]-\int\dd^3x\,(A_V)_iJ_V^i+O(A_V^2)
\]
for a static spatial perturbation.  The commutator kernel and the physical
response kernel then differ by this Hamiltonian sign and by local source
variation terms:
\[
  \Gamma^{ij}(\omega,\mathbf k)
  =
  \frac{\delta\langle J_V^i\rangle}{\delta(A_V)_j}
  =
  -G^R_{J^iJ^j}(\omega,\mathbf k)+K_{\rm loc}^{ij}(\omega,\mathbf k).
\]
Here \(K_{\rm loc}\) contains Bardeen--Zumino and magnetization contact
terms when the chosen current representative includes them.  For
\(\mathbf k=k\hat e_n\),
\[
  B_V^i(\mathbf k)
  =
  \ii\epsilon^{inj}k_n(A_V)_j
  =
  -\ii\epsilon^{ijn}k_n(A_V)_j .
\]
If \(\Gamma_{\rm tr}^{ij}\) is the response kernel after the local
magnetization pieces appropriate to the transport observable have been
removed, the chiral magnetic transport coefficient is
\[
  \xi_B
  =
  \lim_{k\to0}
  \frac{\ii}{2k}
  \epsilon_{ijn}\,
  \Gamma_{\rm tr}^{ij}(\omega=0,\mathbf k=k\hat e_n).
\]
Equivalently, if no local contact kernel is being added to the response,
\[
  \xi_B
  =
  -\lim_{k\to0}
  \frac{\ii}{2k}
  \epsilon_{ijn}\,
  G^R_{J^iJ^j}(\omega=0,\mathbf k=k\hat e_n).
\]
This is a definition in the declared static, homogeneous limit.  If the
limits \(\omega\to0\) and \(k\to0\) fail to commute, the chosen order is part
of the observable.

\section{Equilibrium Derivation}

Place the system on a stationary background
\[
  \dd s^2= -\ee^{2\sigma}(dt+a_i dx^i)^2+h_{ij}dx^i dx^j,
  \qquad
  A=A_0(dt+a_i dx^i)+A_i dx^i .
\]
Gauge and Kaluza--Klein invariant three-dimensional fields are
\[
  \mathcal A_i=A_i-A_0a_i,\qquad A_0 .
\]
An anomalous equilibrium functional contains Chern--Simons-type terms
\[
  W_{\mathrm{odd}}
  =
  \int_{\Sigma_3}
  \left[
    c_1 A_0\,\mathcal A\wedge\dd\mathcal A
    +
    c_2 A_0^2\,\mathcal A\wedge\dd a
    +
    c_3 A_0^3\,a\wedge\dd a
  \right],
\]
with coefficients constrained by the anomaly and by large gauge
transformations.  Differentiating this functional with respect to
\(\mathcal A_i\) and \(a_i\) gives the parity-odd equilibrium pieces of
\(J^i\) and \(T^{0i}\).  The derivation is algebraic once the equilibrium
functional and its gauge-variation convention are fixed.

\begin{remark}[Hydrostatic Chern--Simons variation]
\label{rem:hydrostatic-cs-variation}
Let \(\Sigma_3\) be closed, and let \(A\) and \(a\) be ordinary spatial
one-forms on \(\Sigma_3\).  For constant coefficients \(c_{AA}\), \(c_{Aa}\),
and \(c_{aa}\), set
\begin{equation}
  W_{\rm CS}[A,a]
  =
  \int_{\Sigma_3}
  \left[
    c_{AA} A\wedge\dd A
    +c_{Aa} A\wedge\dd a
    +c_{aa} a\wedge\dd a
  \right].
  \label{eq:hydrostatic-cs-functional-general}
\end{equation}
In local coordinates with orientation
\(\dd^3x\) and \(\epsilon^{123}=1\), define
\begin{equation}
  B_A^i=\epsilon^{ijk}\partial_j A_k,\qquad
  B_a^i=\epsilon^{ijk}\partial_j a_k .
  \label{eq:hydrostatic-magnetic-vectors}
\end{equation}
Then
\begin{equation}
  \frac{\delta W_{\rm CS}}{\delta A_i}
  =
  2c_{AA}B_A^i+c_{Aa}B_a^i,
  \qquad
  \frac{\delta W_{\rm CS}}{\delta a_i}
  =
  c_{Aa}B_A^i+2c_{aa}B_a^i .
  \label{eq:hydrostatic-cs-variation-formula}
\end{equation}
If the coefficients are functions of \(A_0\) or of thermodynamic scalars,
\eqref{eq:hydrostatic-cs-variation-formula} is the part of the variation
with respect to the spatial one-forms at fixed scalar sources; variations of
the scalar sources produce the corresponding charge and energy densities.

On a closed \(\Sigma_3\),
\[
  \delta\int A\wedge\dd A
  =
  \int\delta A\wedge\dd A+\int A\wedge\dd(\delta A)
  =
  2\int\delta A\wedge\dd A,
\]
because \(0=\int\dd(A\wedge\delta A)
=\int\dd A\wedge\delta A-\int A\wedge\dd(\delta A)\) and
\(\dd A\wedge\delta A=\delta A\wedge\dd A\).  Similarly,
\[
  \delta\int A\wedge\dd a
  =
  \int\delta A\wedge\dd a+\int \dd A\wedge\delta a,
  \qquad
  \delta\int a\wedge\dd a
  =
  2\int\delta a\wedge\dd a .
\]
Writing
\(\delta A\wedge\dd A=\delta A_i\epsilon^{ijk}\partial_jA_k\,\dd^3x\)
and the analogous formulas for \(a\) gives
\eqref{eq:hydrostatic-cs-variation-formula}.
\end{remark}

\section{Chiral Magnetic Coefficient}

There are two axial parameters in common chiral-magnetic formulas, and they
must not be merged.  The first is the axial background source
\[
  a_0^5=(A_A)_0 .
\]
The second is the occupation imbalance
\[
  \mu_5^{\rm occ}=\frac{\mu_R-\mu_L}{2}
\]
appearing in a density matrix.  The primary CME protocol in this section is
a prepared, slowly relaxing chiral imbalance,
\begin{equation}
  \rho_{\rm prep}
  =
  Z^{-1}\exp\{-\beta(H-\mu_VQ_V-\mu_5^{\rm occ}Q_5)\},
  \label{eq:cme-prepared-density-matrix}
\end{equation}
used either in a source sector with \(F_V\wedge F_V=0\) during preparation
or on time scales shorter than the anomaly-relaxation time.  The axial
source ledger is kept separately because a vector-preserving equilibrium
functional with \(a_0^5\ne0\) computes a local source current, not the same
density matrix.

First fix the source ledger.  Hold the axial source constant in Euclidean
time,
\[
  (A_A)_0=a_0^5,\qquad \partial_i a_0^5=0,
\]
and set the spatial axial field to zero.  The inflow functional for the
vector-vector-axial anomaly contains the five-dimensional local term
\[
  W_5[A_A,A_V]
  =
  -\frac1{4\pi^2}
  \int_{X_5} A_A\wedge F_V\wedge F_V ,
\]
whose axial gauge variation is
\[
  \delta_\beta W_5
  =
  -\frac1{4\pi^2}\int_{\partial X_5}\beta\,F_V\wedge F_V .
\]
Since
\[
  F_V\wedge F_V
  =
  \frac14\epsilon^{\mu\nu\rho\sigma}
  (F_V)_{\mu\nu}(F_V)_{\rho\sigma}\,\dd^4x ,
\]
this is the inflow form of the axial anomaly displayed above, with the sign
fixed by the convention that the combined bulk-boundary system is invariant.
Let \(\beta\) be the circumference of the thermal circle and let
\(W_3=\ln Z_{\rm odd}\) denote the dimensionally reduced Euclidean exponent.
Reducing the same local term on the thermal circle gives
\begin{equation}
  W_{3,{\rm source}}[A_V;a_0^5]
  =
  -\frac{\beta a_0^5}{4\pi^2}
  \int_{\Sigma_3} A_V\wedge\dd A_V .
  \label{eq:cme-source-three-functional}
\end{equation}
In components,
\[
  \int_{\Sigma_3} A_V\wedge\dd A_V
  =
  \int\dd^3x\,
  \epsilon^{ijk}(A_V)_i\partial_j(A_V)_k .
\]
The four-dimensional consistent current is obtained by dividing the
three-dimensional source variation by \(\beta\):
\begin{equation}
  J_{{\rm cons},V}^i\big|_{\rm source}
  =
  \frac1\beta
  \frac{\delta W_{3,{\rm source}}}{\delta (A_V)_i}
  =
  -\frac{a_0^5}{2\pi^2}
  B_V^i .
  \label{eq:cme-source-only-consistent-current}
\end{equation}
The associated Bardeen--Zumino polynomial in this vector-preserving
convention is
\begin{equation}
  J_{{\rm BZ},V}^\mu
  =
  -\frac1{4\pi^2}
  \epsilon^{\mu\nu\rho\sigma}
  (A_A)_\nu(F_V)_{\rho\sigma},
  \label{eq:cme-vector-bz-current}
\end{equation}
so, for the static source above and the spatial orientation used in
\eqref{eq:hydrostatic-magnetic-vectors},
\[
  J_{{\rm BZ},V}^i
  =
  \frac{a_0^5}{2\pi^2}B_V^i .
\]
The source-only covariant local current therefore vanishes:
\begin{equation}
  J_{{\rm cov},V}^i\big|_{\rm source}
  =
  J_{{\rm cons},V}^i\big|_{\rm source}
  +J_{{\rm BZ},V}^i
  =
  0 .
  \label{eq:cme-source-only-covariant-current}
\end{equation}
This cancellation is the reason an axial thermal holonomy and an axial
occupation imbalance cannot be identified by notation alone.

The prepared-imbalance protocol gives the Fermi-surface part.  For
\(a_0^5=0\) and \(B_V=B\hat z\), the lowest Landau level degeneracy per unit
transverse area is \(|B|/(2\pi)\).  The right-handed lowest Landau level has
group velocity along \(B\), while the left-handed lowest Landau level has
group velocity opposite to \(B\).  Filling right- and left-handed modes to
\(\mu_R\) and \(\mu_L\) in the density matrix
\eqref{eq:cme-prepared-density-matrix} gives
\[
  J_R^z=\frac{\mu_R}{4\pi^2}B,\qquad
  J_L^z=-\frac{\mu_L}{4\pi^2}B .
\]
Thus the local vector current carried by the occupied modes is
\begin{equation}
  J_{V,{\rm occ}}^i
  =
  \frac{\mu_5^{\rm occ}}{2\pi^2}B_V^i .
  \label{eq:cme-prepared-imbalance-current}
\end{equation}
With no axial background source, the vector consistent and covariant current
representatives coincide for this CME term.  With both an occupation
imbalance and a constant axial source present, the convention ledger is
\begin{equation}
\begin{aligned}
  J_{{\rm cons},V}^i
  &=
  \frac{\mu_5^{\rm occ}-a_0^5}{2\pi^2}B_V^i,\\
  J_{{\rm BZ},V}^i
  &=
  \frac{a_0^5}{2\pi^2}B_V^i,\\
  J_{{\rm cov},V}^i
  &=
  \frac{\mu_5^{\rm occ}}{2\pi^2}B_V^i .
\end{aligned}
  \label{eq:cme-consistent-covariant-ledger}
\end{equation}
The last line is the covariant local CME current in the prepared-imbalance
protocol.  It is not the variational current of
\eqref{eq:cme-source-three-functional}.

For the physical electromagnetic current of a Dirac fermion of charge
\(Q=eq\), \(B_V=Q B_{\rm em}\) and
\(J_{{\rm em},{\rm occ}}=QJ_{V,{\rm occ}}\), hence
\begin{equation}
  J_{{\rm em},{\rm occ}}^i
  =
  \frac{e^2q^2\mu_5^{\rm occ}}{2\pi^2}B_{\rm em}^i .
  \label{eq:cme-electromagnetic-current}
\end{equation}

A regulator-level free-Dirac computation has the same split.  A
vector-gauge-invariant regulator fixes the counterterm that produces
\eqref{eq:cme-source-three-functional}; the occupied lowest Landau levels
produce the Fermi-surface term \(\mu_5^{\rm occ}B_V^i/(2\pi^2)\); and the
local Bardeen--Zumino polynomial \eqref{eq:cme-vector-bz-current} changes the
current representative without changing the density matrix.  In a finite
volume, the local hydrostatic part may be a magnetization/contact current
whose spatial integral is a boundary datum.  A net transport CME coefficient
therefore requires the Kubo order above and the subtraction of such local
magnetization terms.

\begin{center}
\scriptsize
\setlength{\tabcolsep}{2pt}
\begin{tabular}{@{}>{\raggedright\arraybackslash}p{0.14\textwidth}
  >{\raggedright\arraybackslash}p{0.24\textwidth}
  >{\raggedright\arraybackslash}p{0.18\textwidth}
  >{\raggedright\arraybackslash}p{0.24\textwidth}
  >{\raggedright\arraybackslash}p{0.10\textwidth}@{}}
Protocol & Source and gauge datum & Density matrix charge & Current and Ward identity & State \\
\hline
Axial source only &
\(a_0^5\ne0\), vector gauge transformations exact, axial source held fixed
in the chosen hydrostatic gauge &
No conserved axial charge inserted &
\(J_{\rm cons,V}=-a_0^5B_V/(2\pi^2)\);
\(J_{\rm cov,V}=0\); vector Ward identity exact &
Equil. \\
Prepared imbalance &
\(a_0^5=0\), magnetic vector source external &
\(Q_5\) used while anomaly production is absent or slower than the
measurement &
\(J_{\rm cons,V}=J_{\rm cov,V}=\mu_5^{\rm occ}B_V/(2\pi^2)\);
axial charge not exact in general &
Metastable \\
Modified axial charge &
Vector source exact; boundary and helicity data included in the conserved
combination &
Conserved \(Q_5\) plus Chern--Simons/helicity contribution &
Observable includes the stored Chern--Simons term; local representative must
be declared &
Equil. + data \\
Transport extraction &
Retarded vector-source perturbation with \(H=H_0-\int A_iJ^i\) &
Whichever state above is being probed &
\(\Gamma_{\rm tr}\) after Bardeen--\allowbreak Zumino and magnetization contact terms
are assigned or subtracted &
Response
\end{tabular}
\end{center}

\section{Chiral Vortical Coefficients}

For one massless Dirac fermion, the covariant first-derivative vortical
currents in the same source-normalized convention are
\begin{align}
  J_V^\mu\big|_\omega
  &=
  \frac{\mu_V\mu_A}{\pi^2}\,\omega^\mu ,
  \label{eq:vector-cve-dirac}\\
  J_A^\mu\big|_\omega
  &=
  \left[
    \frac{\mu_V^2+\mu_A^2}{2\pi^2}
    +\frac{T^2}{6}
  \right]\omega^\mu .
  \label{eq:axial-cve-dirac}
\end{align}
These formulas follow by adding the two Weyl contributions
\[
  J_R^\mu\big|_\omega
  =
  \left(\frac{\mu_R^2}{4\pi^2}+\frac{T^2}{12}\right)\omega^\mu,
  \qquad
  J_L^\mu\big|_\omega
  =
  -\left(\frac{\mu_L^2}{4\pi^2}+\frac{T^2}{12}\right)\omega^\mu .
\]
The vector current is the sum \(J_R+J_L\), while the axial current is the
difference \(J_R-J_L\).  Hence the \(T^2/12\) terms cancel in the vector
current and add in the axial current.  There is therefore no universal
source-normalized vector-current term proportional to \(\mu_A T\) in this
Dirac example; the universal temperature-squared contribution belongs to the
axial vortical current and is tied to the mixed axial-gravitational anomaly.

\section{Entropy Constraint}

Let
\[
  \nabla_\mu T^{\mu\nu}=F^{\nu\lambda}J_\lambda,\qquad
  \nabla_\mu J^\mu=\mathcal A
\]
with anomaly density \(\mathcal A\).  The canonical entropy current is
\[
  S^\mu_{\mathrm{can}}
  =
  p\beta^\mu-\beta_\nu T^{\mu\nu}-\alpha J^\mu,
  \qquad
  \beta^\mu=\frac{u^\mu}{T},\quad \alpha=\frac{\mu}{T}.
\]
The following derivation fixes the signs in this formula.  It is independent
of any derivative expansion.

\paragraph{Canonical entropy divergence with anomaly.}
Let
\begin{equation}
  T^{\mu\nu}
  =
  \epsilon u^\mu u^\nu+p\Delta^{\mu\nu}+\tau^{\mu\nu},
  \qquad
  J^\mu=nu^\mu+\nu^\mu,
  \label{eq:anomalous-transport-dissipative-decomposition}
\end{equation}
with \(u_\mu\tau^{\mu\nu}=0\) and \(u_\mu\nu^\mu=0\).  Assume the
thermodynamic relations
\begin{equation}
  \dd p=s\,\dd T+n\,\dd\mu,\qquad
  \epsilon+p=Ts+\mu n.
  \label{eq:anomalous-transport-thermodynamics}
\end{equation}
Then the canonical entropy current satisfies
\begin{equation}
  \nabla_\mu S^\mu_{\rm can}
  =
  -\tau^{\mu\nu}\nabla_\mu\beta_\nu
  -\nu^\mu\left(\nabla_\mu\alpha-\beta^\nu F_{\mu\nu}\right)
  -\alpha\,\mathcal A .
  \label{eq:canonical-entropy-divergence-anomaly}
\end{equation}

The calculation is as follows.
Differentiate
\[
  S^\mu_{\rm can}=p\beta^\mu-\beta_\nu T^{\mu\nu}-\alpha J^\mu .
\]
Using the anomalous Ward identities gives
\begin{equation}
\begin{aligned}
  \nabla_\mu S^\mu_{\rm can}
  &=
  \beta^\mu\nabla_\mu p
  +p\nabla_\mu\beta^\mu
  -T^{\mu\nu}\nabla_\mu\beta_\nu
  -J^\mu\nabla_\mu\alpha
  -\beta_\nu F^{\nu\lambda}J_\lambda
  -\alpha\mathcal A .
\end{aligned}
  \label{eq:entropy-divergence-before-ideal-cancel}
\end{equation}
The last source term can be rewritten as
\begin{equation}
  -J^\mu\nabla_\mu\alpha-\beta_\nu F^{\nu\lambda}J_\lambda
  =
  -J^\mu\left(\nabla_\mu\alpha-\beta^\nu F_{\mu\nu}\right),
  \label{eq:entropy-force-combination}
\end{equation}
using antisymmetry of \(F_{\mu\nu}\).  It remains to show that the ideal
part cancels.  Substituting
\[
  T^{\mu\nu}_{\rm id}=\epsilon u^\mu u^\nu+p(g^{\mu\nu}+u^\mu u^\nu),
  \qquad J^\mu_{\rm id}=nu^\mu,
\]
and \(\beta^\mu=u^\mu/T\), \(\alpha=\mu/T\), the ideal contribution to
\eqref{eq:entropy-divergence-before-ideal-cancel} is
\begin{equation}
  \frac{u^\mu\nabla_\mu p}{T}
  +p\nabla_\mu\left(\frac{u^\mu}{T}\right)
  -T^{\mu\nu}_{\rm id}\nabla_\mu\left(\frac{u_\nu}{T}\right)
  -nu^\mu\nabla_\mu\left(\frac{\mu}{T}\right).
  \label{eq:entropy-ideal-cancellation-line}
\end{equation}
Since \(u^\nu u_\nu=-1\), \(u^\nu\nabla_\mu u_\nu=0\).  A direct expansion
of \eqref{eq:entropy-ideal-cancellation-line} reduces it to
\[
  \frac{u^\mu}{T}
  \bigl(\nabla_\mu p-s\nabla_\mu T-n\nabla_\mu\mu\bigr),
\]
which vanishes by \(\dd p=s\,\dd T+n\,\dd\mu\).  The remaining terms are
exactly those involving \(\tau^{\mu\nu}\), \(\nu^\mu\), and
\(\mathcal A\), giving
\eqref{eq:canonical-entropy-divergence-anomaly}.

Equation~\eqref{eq:canonical-entropy-divergence-anomaly} explains how
anomaly-induced transport can be nondissipative.  The anomaly contributes
the fixed term \(-\alpha\mathcal A\); parity-odd additions to the entropy
current, built from \(B^\mu\), \(\omega^\mu\), and thermodynamic scalars, can
move this term into a total divergence and thereby relate
\(\xi_B,\xi_\omega\) to anomaly coefficients.  Coefficients invisible to
this entropy analysis must instead be fixed by the hydrostatic functional or
by Kubo formulae with the current representative and contact terms specified.

\section{Topological Terms and Contact Terms}

Equilibrium Chern--Simons terms may change local current correlators by
contact terms.  A transport coefficient must therefore state whether the
current is covariant or consistent, which background functional is used, and
which contact terms are included in the retarded correlator.  The physically
measured response is attached to the source protocol and to the contact-term
convention, as well as to the local operator symbol.

\begin{openproblem}[Nonperturbative anomalous transport]
\label{op:nonperturbative-anomalous-transport}
Derive anomaly-induced transport coefficients in interacting QFT directly
from nonperturbative KMS states, local current algebras, Ward identities,
contact terms, and hydrodynamic scaling limits, with the order of limits and
source protocol stated as part of the theorem.
\end{openproblem}
