\chapter{Thermal Gauge Theory And Screening}
\label{ch:thermal-gauge-theory-screening}

\ConventionsMixed

This chapter studies gauge theories at nonzero temperature through static
correlators and their long-distance screening behavior.  Let \(d\) denote
the number of spatial dimensions, so the thermal Euclidean spacetime is
\(S^1_\beta\times\mathbb R^d\) before spatial infrared regularization.  The
coupling and trace conventions are those of the gauge chapters: the
Yang--Mills action is written with
\[
  \frac{1}{4g^2}\operatorname{tr}(F_{\mu\nu}F^{\mu\nu}),
\]
and group invariants are computed using the same invariant bilinear form.

\section{Static Correlators}

Let the Euclidean time circle have length \(\beta=1/T\).  A bosonic field has
Matsubara frequencies \(\omega_n=2\pi nT\).  Static screening is controlled by
the \(n=0\) sector of gauge-invariant correlators.  For an operator
\(\mathcal O(\tau,\mathbf x)\), define its static projection by
\[
  \mathcal O_0(\mathbf x)=T\int_0^\beta d\tau\,\mathcal O(\tau,\mathbf x).
\]
The screening masses are inverse correlation lengths extracted from
\[
  \left\langle \mathcal O_0(\mathbf x)\mathcal O_0(0)\right\rangle_c
  \sim
  \sum_\alpha C_\alpha
  \frac{e^{-M_\alpha |\mathbf x|}}{|\mathbf x|^{(d-1)/2}}
\]
when the static transfer matrix has isolated lowest eigenvalues in the chosen
channel.

The power of \(|\mathbf x|\) in this formula is fixed by the continuum
spatial Fourier transform, not by a convention for the operator.  If the
static momentum-space correlator has an isolated scalar pole
\[
  \widetilde C(\mathbf k)
  =
  \frac{Z}{\mathbf k^2+M^2}
  +\text{terms less singular at }\mathbf k^2=-M^2 ,
\]
then
\begin{equation}
  C(\mathbf x)
  =
  Z\,\frac{M^{d/2-1}}{(2\pi)^{d/2}|\mathbf x|^{d/2-1}}\,
  K_{d/2-1}(M|\mathbf x|)
  +\cdots ,
  \label{eq:thermal-screening-yukawa-bessel}
\end{equation}
and the large-distance asymptotic of the modified Bessel function gives
\[
  C(\mathbf x)
  \sim
  Z\,\frac{M^{(d-3)/2}}{2(2\pi)^{(d-1)/2}}\,
  \frac{e^{-M|\mathbf x|}}{|\mathbf x|^{(d-1)/2}} .
\]
If instead one projects to zero momentum in the transverse directions and
keeps a single separation coordinate \(z\), the same pole gives
\[
  C_{\rm proj}(z)
  =
  \int d^{d-1}x_\perp\,C(z,\mathbf x_\perp)
  =
  \frac{Z}{2M}e^{-M|z|}
  +\cdots .
\]
This projected form is the one most directly read from a spatial transfer
matrix.  With spatial volume first regulated in a box transverse to the
\(z\)-direction, reflection positivity in \(z\) gives a Hilbert space on the
slice \(S^1_\beta\times\mathbb T^{d-1}\), a positive transfer Hamiltonian
\(H_z\), and
\[
  C_{\rm proj}(z)
  =
  \sum_{\alpha\ne0}
  |\langle\Omega_z,\widehat{\mathcal O}_0\alpha\rangle|^2
  e^{-M_\alpha |z|},
  \qquad
  M_\alpha=E_\alpha^{(z)}-E_0^{(z)} .
\]
The sum is over states in the static, gauge-invariant channel created by
\(\mathcal O_0\).  In infinite transverse volume the isolated eigenvalues may
become thresholds, but the leading exponential rate is still a property of
gauge-invariant static correlators, not of gauge-fixed propagator poles.

\section{Electric Screening}

In a weakly coupled high-temperature regime, the temporal gauge field \(A_0\)
behaves in the static sector as an adjoint scalar.  The electric screening
mass is defined perturbatively by the pole of the static longitudinal
propagator, or nonperturbatively by the lowest gauge-invariant state in the
electric channel after the relevant operator has been specified.

These two definitions should not be conflated.  The perturbative Debye mass
is a matching coefficient of the static effective theory in a specified
gauge, regulator, and operator convention.  A physical screening mass is
extracted from a gauge-invariant static correlator, for example from
operators built from \(A_0\) inside the dimensionally reduced theory or from
renormalized Polyakov loops in the original thermal theory.  At weak coupling
the Debye coefficient controls the electric scale \(gT\) and organizes the
gauge-invariant spectrum, but the spectrum itself is defined by the transfer
matrix construction above.

Choose a basis \(T^a_R\) for each representation \(R\) such that
\[
  \operatorname{tr}_R(T^a_RT^b_R)=T_R\delta^{ab},
  \qquad
  f^{acd}f^{bcd}=C_A\delta^{ab}.
\]
With the same coupling \(g\) in the covariant derivative and in the kinetic
term convention above, the one-loop static polarization gives
\begin{equation}
  m_D^2
  =
  g^2T^2\left(
  \frac{C_A}{3}
  +\frac{1}{3}\sum_{\text{Dirac }f}T_{R_f}
  +\frac{1}{3}\sum_{\text{complex }s}T_{R_s}
  +\frac{1}{6}\sum_{\text{real }r}T_{R_r}
  \right).
\label{eq:one-loop-debye-mass-general}
\end{equation}
For \(SU(N)\) with the fundamental trace convention
\(\operatorname{tr}_{\mathbf N}(T^aT^b)=\delta^{ab}\), this means
\(C_A=2N\) and \(T_{\mathbf N}=1\).

\paragraph{Thermal susceptibility integrals.}
For \(n_B(x)=(e^x-1)^{-1}\) and \(n_F(x)=(e^x+1)^{-1}\),
\begin{align}
  \frac1T\int\frac{\dd^3p}{(2\pi)^3}
  n_B(|p|/T)(1+n_B(|p|/T))
  &=
  \frac{T^2}{6},
  \label{eq:thermal-boson-susceptibility-integral}\\
  \frac1T\int\frac{\dd^3p}{(2\pi)^3}
  n_F(|p|/T)(1-n_F(|p|/T))
  &=
  \frac{T^2}{12}.
  \label{eq:thermal-fermion-susceptibility-integral}
\end{align}

Set \(x=|p|/T\).  Then
\[
  \frac1T\int\frac{\dd^3p}{(2\pi)^3}
  n(1\pm n)
  =
  \frac{T^2}{2\pi^2}
  \int_0^\infty x^2 n(x)(1\pm n(x))\,\dd x .
\]
Since
\[
  n_B(1+n_B)=-\frac{\dd n_B}{\dd x},
  \qquad
  n_F(1-n_F)=-\frac{\dd n_F}{\dd x},
\]
integration by parts gives
\[
  \int_0^\infty x^2n_B(1+n_B)\,\dd x
  =
  2\int_0^\infty \frac{x}{e^x-1}\,\dd x
  =
  \frac{\pi^2}{3},
\]
and
\[
  \int_0^\infty x^2n_F(1-n_F)\,\dd x
  =
  2\int_0^\infty \frac{x}{e^x+1}\,\dd x
  =
  \frac{\pi^2}{6}.
\]
Substitution gives
\eqref{eq:thermal-boson-susceptibility-integral} and
\eqref{eq:thermal-fermion-susceptibility-integral}.

\begin{lemma}[Background-field origin and coefficient of the Debye mass term]
\label{lem:debye-mass-origin}
At one loop, the static quadratic term for \(A_0\) in the dimensionally
reduced action is the zero-frequency, zero-momentum second variation of the
background-field effective action.  If \(S[A_0]\) is the Euclidean action in
a fixed regulator and background-field gauge, then for a constant temporal
background
\begin{equation}
  \left.
  \frac{\partial^2\Gamma[A_0]}{\partial A_0^a\partial A_0^b}
  \right|_{A_0=0}
  =
  \left\langle
  \left.
  \frac{\partial^2S[A_0]}{\partial A_0^a\partial A_0^b}
  \right|_{A_0=0}
  \right\rangle
  -
  \left\langle
  \left.
  \frac{\partial S[A_0]}{\partial A_0^a}
  \right|_{A_0=0}
  \left.
  \frac{\partial S[A_0]}{\partial A_0^b}
  \right|_{A_0=0}
  \right\rangle_c .
  \label{eq:debye-background-second-variation}
\end{equation}
The first term is the contact or seagull contribution.  It vanishes for a
strictly linear fermion coupling, but it is present for scalar and gauge
fields and for the ghost/background-gauge operators.  Evaluating this full
second variation gives \eqref{eq:one-loop-debye-mass-general}.
\end{lemma}

\begin{proof}
Let \(Z[A_0]=\int D\Phi\,\exp(-S[\Phi;A_0])\) and
\(\Gamma[A_0]=-\log Z[A_0]\).  Differentiating \(\Gamma\) twice gives
\eqref{eq:debye-background-second-variation}.  This is the precise object
whose zero-momentum limit is matched to the \(A_0^2\) coefficient in the
static effective theory.  Writing only a connected product of first
variations is correct only for fields whose action is linear in the
background being varied.  A complex scalar, for example, contains
\[
  (D_0\phi)^\dagger D_0\phi
  \supset
  A_0^aA_0^b\,\phi^\dagger T^aT^b\phi ,
\]
and the displayed term contributes to
\(\partial^2S/\partial A_0^a\partial A_0^b\).  In the gauge sector the
background-field vector operator, the gauge-fixing term, and the ghost
operator contribute the analogous contact pieces.  The determinant
calculation below keeps all of them.

For a translation-invariant unbroken thermal state,
\eqref{eq:debye-background-second-variation} is diagonal in color and equals
\(\beta V\,\chi\,\delta^{ab}\).  The coefficient \(\chi\) is the adjoint
electric susceptibility.  The three-dimensional matching convention used
below then sets \(m_D^2=g^2\chi\).

A constant \(A_0=A_0^at^a\) shifts
the Matsubara frequencies exactly as an imaginary color chemical potential.
For a weight \(w\) of a representation \(R\), the quadratic operator depends
on
\[
  \omega_n\longmapsto \omega_n+w(A_0),
\]
for bosons, with the analogous anti-periodic frequencies for fermions.  This
statement can be checked directly from the covariant derivative
\(D_0=\partial_\tau+\operatorname{ad}_{A_0}\), and then extended from a
Cartan-valued \(A_0\) to arbitrary \(A_0\) by \(G\)-invariance of the
quadratic form.  Expanding the shifted determinant is equivalent to
\eqref{eq:debye-background-second-variation}: the second derivative of the
operator is the seagull term, and the product of two first derivatives is the
ordinary insertion term.

For one complex bosonic oscillator of energy \(E=|p|\) and charge \(q\), the
thermal part of the free-energy density is
\[
  T\log\left(1-e^{-\beta(E-\mu q)}\right)
  +
  T\log\left(1-e^{-\beta(E+\mu q)}\right).
\]
Putting \(\mu=i a\) and differentiating twice at \(a=0\) gives
\[
  \left.\frac{\partial^2 f_B(i a)}{\partial a^2}\right|_{a=0}
  =
  \frac{2q^2}{T}\,n_B(E/T)(1+n_B(E/T)).
\]
Thus a real bosonic Gaussian determinant, which is one half of the complex
determinant in a real representation, contributes
\[
  \frac{q^2}{T}\,n_B(E/T)(1+n_B(E/T)).
\]
For one Dirac oscillator, the two spin polarizations and the
particle-antiparticle pair give
\[
  \left.\frac{\partial^2 f_F(i a)}{\partial a^2}\right|_{a=0}
  =
  \frac{4q^2}{T}\,n_F(E/T)(1-n_F(E/T)).
\]
After summing over weights, \(q^2\) is replaced by
\(w^aw^bA_0^aA_0^b\), and
\(\sum_w w^aw^b=\operatorname{tr}_R(T^a_RT^b_R)=T_R\delta^{ab}\).

A single real bosonic degree of freedom in a real representation therefore
contributes
\begin{equation}
  \chi_{\rm real\ boson}^{ab}
  =
  T_R\delta^{ab}
  \frac1T\int\frac{\dd^3p}{(2\pi)^3}
  n_B(1+n_B)
  =
  \frac{T^2}{6}T_R\delta^{ab}.
  \label{eq:debye-real-boson-contribution}
\end{equation}
A complex scalar has two real thermal degrees of freedom in \(R\), and hence
contributes \(T^2T_R\delta^{ab}/3\).  A Dirac fermion has two spin
polarizations and particle/antiparticle excitations.  The anti-periodic
Matsubara sum gives the fermionic integral in
the susceptibility calculation above, so
\begin{equation}
  \chi_{\rm Dirac}^{ab}
  =
  4T_R\delta^{ab}
  \frac1T\int\frac{\dd^3p}{(2\pi)^3}
  n_F(1-n_F)
  =
  \frac{T^2}{3}T_R\delta^{ab}.
  \label{eq:debye-dirac-contribution}
\end{equation}
For gauge fields, use background-field gauge around the flat background
\(A_0\).  Since the background field strength vanishes, the quadratic vector
operator is \(-D^2\delta_{\mu\nu}\) and the ghost operator is \(-D^2\), both
in the adjoint representation.  The one-loop contribution to the thermal
free energy is therefore
\[
  \frac12\operatorname{Tr}_{\rm adj,\,vec}\log(-D^2)
  -\operatorname{Tr}_{\rm adj,\,gh}\log(-D^2)
  =
  \left(\frac{4}{2}-1\right)\operatorname{Tr}_{\rm adj}\log(-D^2)
\]
in four spacetime dimensions.  The right side is the determinant of two real
periodic bosonic thermal degrees of freedom in the adjoint representation:
the temporal and longitudinal gauge-dependent pieces have cancelled against
the periodic complex ghost pair.  Using \(T_{\rm adj}=C_A\), this gives
\begin{equation}
  \chi_{\rm gauge}^{ab}
  =
  2\frac{T^2}{6}C_A\delta^{ab}
  =
  \frac{T^2}{3}C_A\delta^{ab}.
  \label{eq:debye-gauge-contribution}
\end{equation}

The four-dimensional one-loop free-energy density contains
\(\frac12\chi^{ab}A_0^aA_0^b\).  Multiplying by the thermal circle length
\(\beta=1/T\) gives the three-dimensional effective action.  Since the
static action is written as
\[
  \int\dd^3x\,
  \frac{m_D^2}{2g_3^2}\operatorname{tr}(A_0^2),
  \qquad
  g_3^2=g^2T,
\]
and \(\operatorname{tr}(A_0^2)=\delta^{ab}A_0^aA_0^b\) in the default trace
convention, matching gives \(m_D^2=g^2\chi\).  Adding the gauge, Dirac,
complex-scalar, and real-scalar susceptibilities gives
\eqref{eq:one-loop-debye-mass-general}.
\end{proof}

\begin{controlledapproximation}[HTL bridge from Debye matching to response]
\label{ca:thermal-screening-htl-bridge}
The Debye coefficient above is the static face of a real-time soft response.
The hard-thermal-loop approximation is the controlled leading response when
the external frequency and momentum satisfy
\[
  \omega,\ |\mathbf k|=O(gT)
\]
while the particles carrying the loop momentum have
\[
  p=O(T).
\]
The calculation keeps the collisionless angular motion of the hard particles
and expands only in the ratio of soft external scale to hard momentum.  It is
therefore not a local Lorentz-invariant mass term for the gauge field, and it
does not determine the nonperturbative magnetic scale.
\end{controlledapproximation}

\begin{lemma}[Static Debye term as the zero-frequency HTL limit]
\label{lem:thermal-screening-htl-static-limit}
In the rest frame of the equilibrium state, let
\[
  v^\mu=(1,\mathbf v),\qquad |\mathbf v|=1,\qquad
  \int_{\mathbf v}(\cdots)
  =
  \int_{S^2}\frac{\dd\Omega_{\mathbf v}}{4\pi}(\cdots).
\]
With mostly-plus Lorentzian convention and
\[
  K_\mu=(-\omega,\mathbf k),
\]
the retarded HTL induced-current kernel
\begin{equation}
  \mathcal K_R^{\mu\nu,ab}(\omega,\mathbf k)
  =
  \delta^{ab}m_D^2
  \left[
    \delta^\mu_0\delta^\nu_0
    -
    \omega\int_{\mathbf v}
    \frac{v^\mu v^\nu}
    {\omega-\mathbf v\cdot\mathbf k+i0}
  \right]
  \label{eq:thermal-screening-retarded-htl-kernel}
\end{equation}
is transverse,
\[
  K_\mu\mathcal K_R^{\mu\nu,ab}=0,
\]
and its static response is
\[
  \mathcal K_R^{00,ab}(0,\mathbf k)=\delta^{ab}m_D^2,\qquad
  \mathcal K_R^{0i,ab}(0,\mathbf k)
  =
  \mathcal K_R^{ij,ab}(0,\mathbf k)=0 .
\]
Thus the one-loop coefficient in
\eqref{eq:one-loop-debye-mass-general} is the zero-frequency electric
susceptibility of a conserved retarded response, not a gauge-field mass
assigned independently of the response kernel.
\end{lemma}

\begin{proof}
The mechanism is most transparent in the gauge-covariant kinetic form.  Let
\(W(x,\mathbf v)\) be an adjoint auxiliary field and set
\[
  (v\cdot D)W(x,\mathbf v)=v^iF_{i0}(x),\qquad
  j^\mu(x)=m_D^2\int_{\mathbf v}v^\mu W(x,\mathbf v).
\]
The current is covariantly conserved before choosing a gauge:
\[
  D_\mu j^\mu
  =
  m_D^2\int_{\mathbf v}(v\cdot D)W
  =
  m_D^2\int_{\mathbf v}v^iF_{i0}
  =
  0,
\]
because \(F_{i0}\) is independent of \(\mathbf v\) and
\(\int_{\mathbf v}v^i=0\).

Linearize around \(A_\mu=0\) and take fields proportional to
\(\exp(-i\omega t+i\mathbf k\cdot\mathbf x)\).  The retarded solution of
\((\partial_t+\mathbf v\cdot\nabla)W=v^iF_{i0}\) is
\[
  W(K,\mathbf v)
  =
  -\,\frac{\mathbf v\cdot\mathbf k\,A_0(K)
  +\omega\,v^iA_i(K)}
  {\omega-\mathbf v\cdot\mathbf k+i0},
\]
where \(F_{i0}=\partial_iA_0-\partial_0A_i\).  Substitution in
\(j^\mu=m_D^2\int_{\mathbf v}v^\mu W\) gives
\[
  j^\mu
  =
  m_D^2\left[
    \delta^\mu_0\delta^\nu_0
    -
    \omega\int_{\mathbf v}
    \frac{v^\mu v^\nu}
    {\omega-\mathbf v\cdot\mathbf k+i0}
  \right]A_\nu .
\]
The \(A_0\) coefficient uses the identity
\[
  -\,\frac{\mathbf v\cdot\mathbf k}
  {\omega-\mathbf v\cdot\mathbf k+i0}
  =
  1-\frac{\omega}{\omega-\mathbf v\cdot\mathbf k+i0}
\]
inside the angular integral and
\(\int_{\mathbf v}v^\mu=\delta^\mu_0\).  Restoring the color factor gives
\eqref{eq:thermal-screening-retarded-htl-kernel}.

Transversality follows directly.  Since
\[
  K_\mu v^\mu=-(\omega-\mathbf v\cdot\mathbf k),
\]
contracting \(K_\mu\) into
\eqref{eq:thermal-screening-retarded-htl-kernel} cancels the denominator and
leaves
\[
  K_\mu\mathcal K_R^{\mu\nu,ab}
  =
  \delta^{ab}m_D^2
  \left[
    -\omega\delta^\nu_0
    +\omega\int_{\mathbf v}v^\nu
  \right].
\]
The angular integral equals \(1\) for \(\nu=0\) and \(0\) for each spatial
index, so the contraction vanishes.

Finally set \(\omega=0\) in the retarded static response at fixed
\(\mathbf k\).  The angular term in
\eqref{eq:thermal-screening-retarded-htl-kernel} carries an explicit factor
of \(\omega\), and the remaining tensor is
\(\delta^{ab}m_D^2\delta^\mu_0\delta^\nu_0\).  This is exactly the electric
mass coefficient matched into the static effective theory.
\end{proof}

\section{Magnetic Sector}

The spatial gauge fields \(A_i\) have no gauge-invariant local mass term in
the dimensionally reduced action.  The high-temperature static magnetic
sector is therefore a three-dimensional Yang--Mills theory with coupling
\[
  g_3^2=g^2T
\]
in four spacetime dimensions.  Its infrared scale is of order \(g_3^2\), and
the corresponding magnetic screening lengths are nonperturbative objects of
the three-dimensional gauge theory.  Perturbative expansions of static
magnetic quantities encounter this scale because massless spatial gauge
fields remain in the reduced action.

\section{Polyakov Loop And Center Symmetry}

The thermal Wilson line at spatial point \(\mathbf x\) is
\[
  P(\mathbf x)=
  \mathcal P\exp\!\left(\ii\int_0^\beta d\tau\,A_0(\tau,\mathbf x)\right).
\]
In a pure gauge theory, its trace in a representation charged under the center
transforms under the center one-form symmetry wrapped on the thermal circle.
The free energy of an external static charge is defined by
\[
  e^{-\beta F_Q}=\left\langle \operatorname{tr}_R P(\mathbf x)\right\rangle
\]
after renormalizing the line and after specifying the state or center-sector
limit.  The separately traced pair
\[
  \left\langle
  \operatorname{tr}_R P_R(\mathbf x)\,
  \operatorname{tr}_{\bar R}P_{\bar R}(\mathbf y)
  \right\rangle
\]
is a gauge-invariant color-averaged static source-pair observable.  It is not
a projection onto a pure singlet channel.  A singlet coordinate requires extra
structure, for example a gauge-fixed definition such as
\[
  d_R^{-1}\left\langle
  \operatorname{tr}_R(P_R(\mathbf x)P_R^\dagger(\mathbf y))
  \right\rangle_{\rm gf},
\]
whose gauge dependence is part of the datum, or a gauge-invariant cyclic
Wilson loop
\[
  d_R^{-1}\left\langle
  \operatorname{tr}_R\!\left(
    P_R(\mathbf x)U_R(\mathbf x,\mathbf y)
    P_R^\dagger(\mathbf y)U_R(\mathbf y,\mathbf x)
  \right)\right\rangle ,
\]
whose spatial transporters introduce line, cusp/intersection, and mixing
renormalizations.  The color-averaged correlator, the gauge-fixed singlet
coordinate, and the cyclic Wilson loop therefore have different
renormalization data and should not be given the same free-energy label.
Their large-distance decays belong to the static screening spectrum rather
than to a zero-temperature particle spectrum.

In a perturbative channel decomposition for a fundamental
\(SU(N)\) source pair, after a singlet/adjoint channel convention has been
supplied, the color-averaged object has the representation
\[
  e^{-\beta F_{\rm avg}(r)}
  =
  \frac1{N^2}e^{-\beta F_1(r)}
  +
  \frac{N^2-1}{N^2}e^{-\beta F_{\rm adj}(r)}.
\]
The weights are only the finite color multiplicities in
\(\mathbf N\otimes\overline{\mathbf N}=\mathbf 1\oplus\mathrm{adj}\); they do
not turn the traced pair itself into a singlet observable.

\begin{definition}[Renormalized static-source free-energy datum]
\label{def:thermal-static-source-free-energy-datum}
At finite cutoff \(\Lambda\), let \(P_R^\Lambda(\mathbf x)\) be the
regularized thermal line in representation \(R\).  A static-source
renormalization scheme consists of additive line energies
\(\delta m_R(\Lambda)\), equivalently multiplicative factors
\[
  Z_R(\Lambda,\beta)=\exp\{\beta\,\delta m_R(\Lambda)\},
  \qquad
  P_R^{\rm ren}(\mathbf x)=Z_R(\Lambda,\beta)P_R^\Lambda(\mathbf x).
\]
For a conjugate pair at separation \(r=|\mathbf x-\mathbf y|\), define
\[
\begin{aligned}
  e^{-\beta F_R^{\rm ren}}
  &:=
  \left\langle d_R^{-1}\operatorname{tr}_R
  P_R^{\rm ren}(\mathbf x)\right\rangle,\\
  e^{-\beta F_{R\bar R}^{\rm ren}(r)}
  &:=
  \left\langle d_R^{-2}
  \operatorname{tr}_R P_R^{\rm ren}(\mathbf x)\,
  \operatorname{tr}_{\bar R}P_{\bar R}^{\rm ren}(\mathbf y)
  \right\rangle .
\end{aligned}
\]
The second line is the color-averaged traced-pair free energy in the declared
line scheme.  It is not the cyclic Wilson-loop free energy and not a
gauge-fixed singlet channel unless those additional definitions are supplied.
Changing the finite part of \(\delta m_R\) shifts \(F_R^{\rm ren}\) by an
additive constant.  The datum therefore includes the line scheme whenever an
absolute one-source free energy is quoted.
\end{definition}

\begin{lemma}[Line self-energy cancellation in static-source ratios]
\label{lem:thermal-static-source-line-cancellation}
In the datum of Definition~\ref{def:thermal-static-source-free-energy-datum},
assume the expectation values are taken in a fixed physical state and
center-sector prescription, or in a source-selected infinite-volume limit, for
which the one-source expectations satisfy
\(C_R^{\rm ren}C_{\bar R}^{\rm ren}\ne 0\).  Then
the pair excess free energy
\[
  \Delta F_{R\bar R}(r)
  =
  F_{R\bar R}^{\rm ren}(r)-F_R^{\rm ren}-F_{\bar R}^{\rm ren}
\]
is independent of the multiplicative line factors \(Z_R,Z_{\bar R}\).  So is
the static force
\[
  \mathcal F_{R\bar R}(r,T)=-\partial_rF_{R\bar R}^{\rm ren}(r,T),
\]
which requires no division by the one-source expectations, provided the pair
correlator defines \(F_{R\bar R}^{\rm ren}\) and the line counterterms are
local energies of the individual wrapped sources and hence independent of
\(r\).
\end{lemma}

\begin{proof}
Write the cutoff expectations as
\[
  C_R^\Lambda=\langle P_R^\Lambda\rangle,\qquad
  C_{R\bar R}^\Lambda(r)=
  \langle P_R^\Lambda(\mathbf x)P_{\bar R}^\Lambda(\mathbf y)\rangle
\]
with the displayed normalized traces suppressed.  Renormalization gives
\[
  C_R^{\rm ren}=Z_R C_R^\Lambda,\qquad
  C_{R\bar R}^{\rm ren}(r)=Z_RZ_{\bar R}C_{R\bar R}^\Lambda(r).
\]
When \(C_R^\Lambda C_{\bar R}^\Lambda\ne0\), equivalently after finite
renormalization \(C_R^{\rm ren}C_{\bar R}^{\rm ren}\ne0\), therefore
\[
  \exp\{-\beta\Delta F_{R\bar R}(r)\}
  =
  \frac{C_{R\bar R}^{\rm ren}(r)}
       {C_R^{\rm ren}C_{\bar R}^{\rm ren}}
  =
  \frac{C_{R\bar R}^{\Lambda}(r)}
       {C_R^\Lambda C_{\bar R}^\Lambda},
\]
so all one-line factors cancel before the continuum limit is discussed.  This
is a conditional ratio statement, not a definition in a sector where either
one-point function vanishes.  For the force, no one-point divisor is used:
\[
  F_{R\bar R}^{\rm ren}(r,T)
  =
  F_{R\bar R}^{\Lambda}(r,T)
  -\delta m_R(\Lambda)-\delta m_{\bar R}(\Lambda),
\]
and differentiating with respect to \(r\) removes the same local constants.
\end{proof}

Thus the absolute number assigned to a single static-source free energy is a
line-scheme coordinate, while the conditional connected Polyakov-loop ratio
and the source-pair force are the quantities that can be compared across
regulators after the same physical state and center-sector limit have been
specified.  In a center-symmetric finite-volume pure gauge theory the
nonzero-\(N\)-ality one-point function still vanishes before any finite
multiplicative renormalization; the source-selected infinite-volume limit is
the separate order-parameter limit used in phase discussions.  In the
center-symmetric sector itself one instead uses the neutral pair correlator
\[
  G_{R\bar R}^{\rm ren}(r)
  =
  \left\langle d_R^{-2}
  \operatorname{tr}_R P_R^{\rm ren}(\mathbf x)\,
  \operatorname{tr}_{\bar R}P_{\bar R}^{\rm ren}(\mathbf y)
  \right\rangle
\]
or, where this correlator is nonzero and differentiable in \(r\), the direct
pair force
\[
  -\partial_rF_{R\bar R}^{\rm ren}(r,T)
  =
  \beta^{-1}\partial_r\log G_{R\bar R}^{\rm ren}(r),
\]
without assigning finite one-source free energies.

\section{Dimensional Reduction}

\begin{controlledapproximation}[Local EQCD matching]
\label{ca:thermal-eqcd-local-matching}
Local electrostatic QCD is a high-temperature weak-coupling expansion, not a
formal consequence of taking \(|\mathbf x|\gg\beta\).  Work in a fixed
gauge-invariant ultraviolet regulator and match static correlation functions
or a background-field effective action in a specified scheme.  Assume a
scale hierarchy
\begin{equation}
  |\mathbf p|,\ m_E,\ g_3^2 \ll M_{\rm hard},
  \qquad
  M_{\rm hard}\sim 2\pi T,
  \qquad
  g(2\pi T)\ll1,
  \label{eq:thermal-eqcd-scale-hierarchy}
\end{equation}
where \(M_{\rm hard}\) is a lower bound on the modes removed from the
effective theory.  In the ordinary deconfined weak-coupling expansion those
removed modes are the nonzero bosonic Matsubara modes and all thermal
fermion modes; the retained light fields are the static spatial gauge field
and the adjoint scalar \(A_0\).  The exact static functional obtained by
integrating out the removed modes can be nonlocal.  The local EQCD action is
the Taylor expansion of that functional in external static momenta and light
fields, with a declared operator basis and a truncation order.

In four spacetime dimensions the leading bosonic action begins as
\[
  S_{\mathrm{EQCD}}
  =
  \int d^3x\,
  \left[
  \frac{1}{4g_3^2}\operatorname{tr}(F_{ij}F_{ij})
  +\frac{1}{2g_3^2}\operatorname{tr}(D_iA_0D_iA_0)
  +\frac{m_E^2}{2g_3^2}\operatorname{tr}(A_0^2)
  +\lambda_E\,\operatorname{tr}(A_0^4)+\cdots
  \right].
\]
The coefficients \(g_3^2,m_E^2,\lambda_E,\ldots\) are fixed by matching in
the same regulator and operator scheme as the original thermal theory, as in
the dimensional-reduction matching programs of
\cite{BraatenNieto1995EFTThermo,KajantieEtAl1996DimensionalReduction}.  If the
operator basis is kept through order \(D_\ast\), the omitted terms are
controlled only in the matching domain.  Their contribution to a static
amplitude at momentum \(p\) is bounded, in the declared matching expansion, by
powers of
\begin{equation}
  \frac{p}{M_{\rm hard}},\qquad
  \frac{m_E}{M_{\rm hard}},\qquad
  \frac{g_3^2}{M_{\rm hard}},
  \label{eq:thermal-eqcd-expansion-parameters}
\end{equation}
with scheme-dependent coefficients.  The magnetic sector of the
three-dimensional theory still contains nonperturbative physics at the
\(g_3^2\) scale.  EQCD locality does not remove that infrared problem.
\end{controlledapproximation}

The hypothesis \(M_{\rm hard}\sim2\pi T\) includes a holonomy condition.  In
a static diagonal background with eigenvalue differences \(\alpha(A_0)\), an
adjoint mode has covariant Matsubara frequency
\[
  2\pi nT+\alpha(A_0).
\]
If this quantity is small for some mode that was going to be integrated out,
that mode is not hard and must either be retained or treated by a different
effective description.  The same warning applies near thermal transitions or
strong-coupling regimes where \(g(2\pi T)\) is not a small matching parameter,
to backgrounds with imaginary chemical potentials that shift fermionic
thermal gaps, and to theories with additional light scalars.  In those cases
the exact static functional may still exist, but the local EQCD truncation
above has lost the scale separation that justified it.

This separates three objects that often appear in the same sentence.  The
perturbative Debye parameter \(m_E\) is a matched coefficient in
Controlled Approximation~\ref{ca:thermal-eqcd-local-matching}.  The
gauge-invariant screening spectrum is extracted from static transfer-matrix
correlators.  Static-source free energies are Polyakov-line observables with
their own line, channel, and center-sector data.  Agreement among their
leading weak-coupling scales is useful physics, but it is not an identity of
definitions.

\begin{openproblem}[Gauge-invariant screening spectrum]
\label{op:gauge-invariant-screening-spectrum}
Give a continuum construction of the full static screening spectrum of
thermal non-Abelian gauge theory from gauge-invariant operators, including
the relation between perturbative electric matching, nonperturbative magnetic
dynamics, and Polyakov-loop sectors.
\end{openproblem}
