\chapter{Supersymmetric Localization On Compact Manifolds}
\label{ch:susy-localization-compact-manifolds}

\ConventionsMixed

Supersymmetric localization is an exact-evaluation method only after the
regulated functional integral has been specified with enough information to
justify deformation of the integration cycle.  This chapter separates the
finite-dimensional theorem from the QFT formulae used in four-dimensional
\(\mathcal N=2\) gauge theory on \(S^4\) and three-dimensional
\(\mathcal N=2\) Chern--Simons--matter theory on \(S^3\).  The formulae below
are therefore not introduced as formal replacements for a definition of a QFT:
they are consequences of a localization structure in the sense of
Definition~\ref{def:susy-localization-data}, together with the determinant
and instanton-sector prescriptions explicitly stated here.

\begin{definition}[Compact supersymmetric localization structure]
\label{def:compact-supersymmetric-localization-datum}
A compact supersymmetric localization structure is a tuple
\[
  \mathfrak L_M
  =
  \bigl(
    M,\mathfrak B,
    \{(\mathfrak F_\Lambda,C_\Lambda,\mu_\Lambda)\}_{\Lambda\in I},
    Q_\Lambda,B_\Lambda,V_\Lambda,
    S_\Lambda,\mathcal O_\Lambda,
    \mathfrak F^{\rm fix}_\Lambda,
    \mathcal N^{\bullet}_\Lambda,
    \mathfrak R_\Lambda,
    \mathcal C_\Lambda
  \bigr)
\]
with the following meaning.
\begin{enumerate}[label=(\roman*)]
\item \(M\) is the compact oriented Riemannian manifold on which the QFT is
placed, and \(\mathfrak B\) denotes the chosen rigid-supergravity background:
spin or spin-\(c\) structure, \(R\)-symmetry and flavor bundles, background
connections, masses, and any metric or framing data used by the preserved
supercharge.
\item \(I\) is a directed regulator set.  For each \(\Lambda\in I\),
\(\mathfrak F_\Lambda\) is the regulated super field-variable space, including
gauge-fixing or BV variables when gauge symmetry is present; \(C_\Lambda\) is
the declared integration cycle, including orientation, boundary, singular
strata, and gauge quotient; and \(\mu_\Lambda\) is the regulated Berezinian
density or BV half-density.  Fermionic variables are understood as coordinates
on a locally super-ringed space, so \(\mu_\Lambda\) is not a Borel measure.
\item \(Q_\Lambda\) is an odd derivation of the regulated algebra of
observables, and
\[
  B_\Lambda:=Q_\Lambda^2
  =
  \mathcal L_{v_\Lambda}
  +\delta_{\rm gauge}(\Lambda_Q)
  +\delta_R(\Theta_R)+\delta_F(\Theta_F)
\]
is the specified even transformation: an isometry of \(M\), a gauge
transformation, and \(R\)- or flavor rotations.  In a gauge theory this
statement is made in the BV complex; compatibility means that \(Q_\Lambda\)
commutes with the gauge-fixed BV differential up to \(B_\Lambda\)-symmetry and
that the regulated quantum master equation has the residual boundary terms
specified below.
\item \(S_\Lambda\) is the regulated action and \(\mathcal O_\Lambda\) is the
regulated protected insertion.  They obey
\[
  Q_\Lambda S_\Lambda=0,\qquad
  Q_\Lambda\mathcal O_\Lambda=0,\qquad
  \mathcal L_{Q_\Lambda}\mu_\Lambda=0
\]
in the regulated theory, or else the anomaly term is stated explicitly as a
defect of these equations.
\item \(V_\Lambda\) is an odd localizing functional such that
\(\operatorname{Re}(Q_\Lambda V_\Lambda)\ge 0\) on the bosonic body of
\(C_\Lambda\), with equality precisely on a stated fixed locus
\(\mathfrak F^{\rm fix}_\Lambda\).  The \(Q_\Lambda\)-Stokes boundary term
\[
  \int_{\partial C_\Lambda}
  \mu_\Lambda\,\mathcal O_\Lambda V_\Lambda
  \exp\{-S_\Lambda-tQ_\Lambda V_\Lambda\}
\]
is either proved to vanish, canceled by a declared relative-cycle or boundary
counterterm, or recorded as an open obstruction.
\item \(\mathcal N_\Lambda^\bullet\) is the normal \(Q_\Lambda\)-complex to
the smooth part of \(\mathfrak F^{\rm fix}_\Lambda\), with zero modes split
off as coordinates or insertion modes.  The one-loop determinant is the
Berezin Gaussian/Euler class of the acyclic part of this complex, with
orientation, phase, zeta or heat-kernel regularization, framing, and local
counterterm conventions included in the structure.
\item \(\mathfrak R_\Lambda\) is the residual finite-dimensional integration
or summation prescription after normal integration: Cartan contour, residue
chamber, Jeffrey--Kirwan vector, instanton compactification, polarization,
or equivariant pushforward.  Singular fixed strata such as zero-size
instantons are part of \(\mathfrak R_\Lambda\), not automatic consequences of
the notation \(\int\mathcal D\Phi\).
\item \(\mathcal C_\Lambda\) is the continuum comparison structure: topology on
protected observables, renormalization and contact-term scheme, convergence
or matching statement for the limit \(\Lambda\to\infty\), and the precise
status of any equality between the regulated localization formula and the
underlying QFT observable.
\end{enumerate}
All localization formulae in this chapter are statements about consequences
of a structure \(\mathfrak L_M\).  If one of the entries is supplied only by a
standard prescription rather than by a regulator theorem, the corresponding
formula is a controlled prescription or hypothesis at that entry, even when
the finite-dimensional localization identity itself is a theorem.
\end{definition}

\section{Finite-Dimensional Deformation Identity}

Let \(X\) be an oriented finite-dimensional supermanifold equipped with an
integration cycle \(C\subset X\) and Berezinian density \(\mu\).  Let \(Q\) be
an odd vector field on \(X\), and let \(B\) be the even vector field \(Q^2\).
For the following proposition all functions and densities are assumed smooth
on a neighborhood of \(C\), and \(C\) is assumed to be either compact without
boundary or equipped with boundary conditions for which the displayed
Stokes term vanishes.

\begin{remark}[Localization deformation identity]
\label{rem:susy-localization-deformation-identity}
Let \(S,V,\mathcal O\) be functions on \(X\) with parities
\(|S|=|\mathcal O|=0\), \(|V|=1\).  Suppose
\[
  Q S=0,\qquad Q\mathcal O=0,\qquad
  B S=B\mathcal O=BV=0,
\]
and suppose the Berezinian density is \(Q\)-invariant:
\(\mathcal L_Q\mu=0\).  Define
\[
  I(t)=\int_C\mu\,\mathcal O\,\exp\{-S-t\,QV\}.
\]
If \(Q(\mu\,\mathcal O\,V\,\exp\{-S-tQV\})\) has zero integral over \(C\),
then \(I(t)\) is independent of \(t\).
Because \(Q\) is an odd derivation and \(Q^2V=BV=0\),
\[
  \frac{dI}{dt}
  =
  -\int_C\mu\,\mathcal O\,(QV)\,\exp\{-S-tQV\}.
\]
The hypotheses \(QS=0\), \(Q\mathcal O=0\), and \(Q(QV)=0\) give
\[
\begin{aligned}
  Q\!\left(\mathcal O\,V\,\exp\{-S-tQV\}\right)
  &=
  (Q\mathcal O)V\ee^{-S-tQV}
  +\mathcal O(QV)\ee^{-S-tQV} \\
  &\quad
  -\mathcal O V\,Q(S+tQV)\ee^{-S-tQV} \\
  &=
  \mathcal O(QV)\ee^{-S-tQV}.
\end{aligned}
\]
Using \(\mathcal L_Q\mu=0\), the right hand side integrates to the assumed
zero Stokes term.  Hence \(dI/dt=0\).
\end{remark}

\begin{definition}[Clean fixed locus and normal complex]
\label{def:clean-fixed-locus-normal-complex}
Let \(F\subset C_{\bar 0}\) be the common zero locus of the bosonic part of
\(QV\) on the body of the integration cycle.  The fixed locus is called clean
for the finite-dimensional localization quadruple \((C,\mu,Q,V)\) if \(F\) is a submanifold, if the Hessian of
\(\operatorname{Re}(QV)\) is nondegenerate on a chosen normal bundle
\(N_{F,\bar 0}\), and if the linearization of \(Q\) in the directions normal
to \(F\) gives a finite \(\mathbb Z_2\)-graded complex
\[
  N_F^\bullet:\qquad
  N_{F,\bar 0}\xrightarrow{\,Q_{\rm lin}\,}N_{F,\bar 1}
  \xrightarrow{\,Q_{\rm lin}\,}N_{F,\bar 0}
\]
whose cohomology has been split off as zero-mode coordinates on \(F\).  The
one-loop determinant is the Berezin Gaussian of this acyclic normal complex,
with orientation and square-root branches inherited from the integration
cycle.
\end{definition}

Assume, in addition to the deformation identity in
Remark~\ref{rem:susy-localization-deformation-identity}, that
\(\operatorname{Re}(QV)\ge0\) on the bosonic body of \(C\), that \(F\) is
compact and clean, and that the large-\(t\) integral is dominated by a
tubular neighborhood of \(F\).  In local coordinates \(y\) along \(F\),
bosonic normal coordinates \(x^i\), and fermionic normal coordinates
\(\vartheta^A\), the even quadratic normal model has the form
\begin{equation}
  (QV)_{\rm quad}
  =
  \frac12 x^i A_{ij}(y)x^j
  -
  \frac12\vartheta^A\mathsf F_{AB}(y)\vartheta^B .
  \label{eq:finite-normal-quadratic-model}
\end{equation}
Here \(A\) is symmetric and positive on the declared real normal cycle, and
\(\mathsf F\) is antisymmetric.  The differential in the normal
\(Q\)-complex is encoded in the fermion-fermion matrix \(\mathsf F\), not in
an odd-even term: an ordinary even quadratic action cannot contain a term
linear in one fermionic normal coordinate and one bosonic normal coordinate
with an even numerical coefficient.  If, after choosing splittings,
\begin{equation}
  \mathsf F=
  \begin{pmatrix}
    0&D\\
    -D^T&0
  \end{pmatrix},
  \label{eq:normal-complex-fermion-block}
\end{equation}
then the corresponding Pfaffian is \(\pm\det D\), with the sign fixed by the
orientation of the odd variables.  With the conventional
finite-dimensional Gaussian normalization absorbed into \(\mu\), the local
normal factor is therefore
\begin{equation}
  Z_{\rm normal}(y)
  =
  \frac{\operatorname{Pf}\mathsf F(y)}
       {\sqrt{\det A(y)}} .
  \label{eq:finite-normal-localization-factor}
\end{equation}

The finite normal Gaussian factor is fixed by the ordinary Gaussian integral
and the defining coefficient of the Pfaffian.  Let \(A\) be an invertible real
symmetric matrix whose real part is positive
on the chosen bosonic normal cycle, and let \(\mathsf F\) be an invertible
antisymmetric \(2m\times2m\) matrix.  With the Berezin orientation
\(d\vartheta_{2m}\cdots d\vartheta_1\),
\begin{equation}
  \int d^{2m}\vartheta\,
  \exp\!\left(\frac12\vartheta^A\mathsf F_{AB}\vartheta^B\right)
  =
  \operatorname{Pf}\mathsf F .
  \label{eq:finite-normal-pfaffian-integral}
\end{equation}
After absorbing the ordinary factor \((2\pi)^{n/2}\) into the local
normalization of the bosonic measure,
\begin{equation}
  \int dx\,d\vartheta\,
  \exp\!\left(
    -\frac12x^iA_{ij}x^j
    +\frac12\vartheta^A\mathsf F_{AB}\vartheta^B
  \right)
  =
  \frac{\operatorname{Pf}\mathsf F}{\sqrt{\det A}} .
  \label{eq:finite-normal-gaussian-integral}
\end{equation}
If \(A\) or \(\mathsf F\) has a zero mode, that direction must be split off
and integrated on the fixed locus or saturated by insertions; it is not part
of the one-loop determinant.
The Berezin integral extracts the coefficient of the ordered monomial
\(\vartheta^1\cdots\vartheta^{2m}\).  By definition of the Pfaffian,
this coefficient in
\(\exp(\frac12\vartheta^A\mathsf F_{AB}\vartheta^B)\) is
\(\operatorname{Pf}\mathsf F\), proving
\eqref{eq:finite-normal-pfaffian-integral}.  The bosonic Gaussian on the
declared cycle is
\[
  \int_{\mathbb R^n}d^nx\,
  \exp\!\left(-\frac12x^iA_{ij}x^j\right)
  =
  \frac{(2\pi)^{n/2}}{\sqrt{\det A}}
\]
after choosing the square-root branch determined by the cycle.  Multiplying
the bosonic and fermionic factors and absorbing \((2\pi)^{n/2}\) into the
local measure gives \eqref{eq:finite-normal-gaussian-integral}.

If a bosonic zero mode is present, the Gaussian integral diverges or becomes
a distribution along that direction; if a fermionic zero mode is present,
the top Berezin coefficient is absent unless an insertion supplies the
missing odd variable.  Thus zero modes are coordinates on \(F\), insertions
to be saturated, or obstructions showing that the proposed localized
observable vanishes.  A determinant formula without this zero-mode statement
is not a localization theorem.

Equation~\eqref{eq:finite-normal-localization-factor} is the finite model for
the one-loop determinants in the compact-space QFT examples below.

\section{QFT Localization Structure}

For a supersymmetric QFT on a compact Riemannian manifold \(M\), the symbol
\[
  Z_M=\int_{\mathfrak F}\mathcal D\Phi\,\ee^{-S[\Phi]}
\]
is shorthand for the regulated data in
Definition~\ref{def:compact-supersymmetric-localization-datum}.  In the
minimal notation one keeps only
\[
  (\mathfrak F_\Lambda,C_\Lambda,\mu_\Lambda,S_\Lambda,Q_\Lambda,V_\Lambda)
\]
together with the limiting prescription \(\Lambda\to\infty\), but the omitted
entries of \(\mathfrak L_M\) remain logically present.  Thus the localized
expression is justified only when the following conditions are verified or
retained as hypotheses:
\[
\begin{gathered}
  Q_\Lambda S_\Lambda=0,\qquad
  \mathcal L_{Q_\Lambda}\mu_\Lambda=0,\qquad
  Q_\Lambda\mathcal O_\Lambda=0,\\
  \int_{\partial C_\Lambda}
  \mu_\Lambda\,\mathcal O_\Lambda V_\Lambda
  \ee^{-S_\Lambda-tQ_\Lambda V_\Lambda}=0,\\
  \operatorname{Re}(Q_\Lambda V_\Lambda)\ge0
  \quad\hbox{on the bosonic integration cycle}.
\end{gathered}
\]
The fixed locus, zero modes, determinants, and singular strata must then be
included in the limiting prescription.
Noncompact Coulomb directions, Jeffrey--Kirwan residues, and point-instanton
compactifications are not automatic outputs of the symbol \(\int\mathcal
D\Phi\).  They are part of the regulated localization structure.  In the
formulae below this structure will be stated by naming the finite-dimensional
variables, the contour, and the determinant or equivariant-integral prescription.

\section{\texorpdfstring
  {\(S^4\) Localization In Four-Dimensional \(\mathcal N=2\) Gauge Theory}
  {S4 Localization In Four-Dimensional N=2 Gauge Theory}}

Let \(G\) be compact connected, with Lie algebra \(\mathfrak g\), maximal
torus \(T\), Cartan algebra \(\mathfrak t\), Weyl group \(W_G\), and root
system \(\Delta\subset\mathfrak t^\ast\).  The monograph convention for
non-Abelian gauge theory is
\[
  \operatorname{tr}_{\square}(t^at^b)=\delta^{ab},
  \qquad
  S_{\rm YM}=\frac1{4g_{\rm YM}^2}
  \int \operatorname{tr}_{\square}(F_{\mu\nu}F^{\mu\nu}) .
\]
With \((a,a):=\operatorname{tr}_{\square}(a^2)\) for \(a\in\mathfrak t\), the
round \(S^4_r\) localization variable is the constant adjoint scalar
\(a\in\mathfrak t\) on the Coulomb locus.  The classical Gaussian in this
trace-delta convention is
\begin{equation}
  Z_{\rm cl}(a)
  =
  \exp\!\left\{
    -\frac{4\pi^2r^2}{g_{\rm YM}^2}(a,a)
  \right\}.
  \label{eq:s4-localization-classical-gaussian}
\end{equation}
The coefficient equals the familiar half-trace coefficient because
\(t^a=\sqrt2\,T^a\) and \(g_{\rm ht}^2=2g_{\rm YM}^2\), so
\[
  \frac{8\pi^2}{g_{\rm ht}^2}\operatorname{tr}_{\rm ht}(a_{\rm ht}^2)
  =
  \frac{4\pi^2}{g_{\rm YM}^2}(a,a).
\]

The localization structure contains more than the Cartan variable.  An off-shell
\(4D\) \(\mathcal N=2\) vector multiplet on the round sphere consists of a
connection \(A_\mu\), complex adjoint scalars \(\phi,\bar\phi\), gaugini
\(\lambda_A,\bar\lambda_A\), \(A=1,2\), and auxiliary fields
\(D_{AB}=D_{BA}\).  The supercharge is specified by conformal Killing spinors
\((\xi_A,\bar\xi_A)\) and their conformal partners
\((\xi'_A,\bar\xi'_A)\) satisfying
\begin{equation}
  \nabla_\mu\xi_A=\sigma_\mu\bar\xi'_A,\qquad
  \nabla_\mu\bar\xi_A=\bar\sigma_\mu\xi'_A,\qquad
  \nabla_\mu\xi'_A=-\frac1{4r^2}\sigma_\mu\bar\xi_A,\qquad
  \nabla_\mu\bar\xi'_A=-\frac1{4r^2}\bar\sigma_\mu\xi_A .
  \label{eq:s4-conformal-killing-spinor-system}
\end{equation}
In stereographic coordinates \(x^\mu\) from the north pole, with
\(\Omega=(1+|x|^2/(4r^2))^{-1}\), a representative is
\begin{equation}
  \xi_A(x)=\Omega^{1/2}\bigl(\epsilon_A+x^\mu\sigma_\mu\bar\eta_A\bigr),
  \qquad
  \bar\xi_A(x)=\Omega^{1/2}\bigl(\bar\epsilon_A+x^\mu\bar\sigma_\mu\eta_A\bigr),
  \label{eq:s4-cks-stereographic-representative}
\end{equation}
with constant spinors chosen so that the bilinear vector
\[
  v^\mu=\xi^A\sigma^\mu\bar\xi_A
\]
vanishes at the north and south poles and generates the two tangent-plane
rotations used in the \(\Omega\)-background limits.  The exact normalization of
the constant spinors fixes only the scale of the equivariant parameters; the
zero set of \(v\) and the closure of the algebra below are the invariant data.

For the vector multiplet, the fermion variations may be written in the usual
spinor variables as
\begin{align}
  QA_\mu
  &=
  \ii\xi^A\sigma_\mu\bar\lambda_A
  -\ii\bar\xi^A\bar\sigma_\mu\lambda_A,
  \label{eq:s4-vector-susy-A}\\
  Q\phi&=-\ii\xi^A\lambda_A,\qquad
  Q\bar\phi=\ii\bar\xi^A\bar\lambda_A,
  \label{eq:s4-vector-susy-scalars}\\
  Q\lambda_A
  &=
  \frac12\sigma^{\mu\nu}\xi_A F_{\mu\nu}
  +2\sigma^\mu\bar\xi_A D_\mu\phi
  +2\ii\xi_A[\phi,\bar\phi]
  +D_{AB}\xi^B
  +2\phi\,\sigma^\mu\nabla_\mu\bar\xi_A,
  \label{eq:s4-vector-susy-lambda}\\
  Q\bar\lambda_A
  &=
  \frac12\bar\sigma^{\mu\nu}\bar\xi_A F_{\mu\nu}
  -2\bar\sigma^\mu\xi_A D_\mu\bar\phi
  -2\ii\bar\xi_A[\phi,\bar\phi]
  +D_{AB}\bar\xi^B
  -2\bar\phi\,\bar\sigma^\mu\nabla_\mu\xi_A .
  \label{eq:s4-vector-susy-lambdabar}
\end{align}
The variation of \(D_{AB}\) is the covariant fermion derivative fixed by
\(Q^2\)-closure.  It is usually cleaner to pass to the cohomological variables
obtained by contracting the gaugini with the chosen spinors:
\[
  (A,\Phi,\bar\Phi;\Psi,\eta,\chi;H).
\]
Here \(\Psi\) is a one-form, \(\eta\) is a scalar, and \(\chi,H\) are sections
of the self-dual two-form bundle rotated by the \(SU(2)_R\) background.  In
these variables the off-shell transformations are
\begin{equation}
\begin{aligned}
  QA &= \Psi,&
  Q\Psi &= \iota_v F+D_A\Phi,\\
  Q\Phi &= \iota_v\Psi,&
  Q\bar\Phi &= \eta,\\
  Q\eta &= \mathcal L_v^A\bar\Phi+[\Phi,\bar\Phi],&
  Q\chi &= H,\\
  QH &= \mathcal L_v^A\chi+[\Phi,\chi]+\Theta\cdot\chi .
\end{aligned}
  \label{eq:s4-cohomological-q-transformations}
\end{equation}
Thus
\begin{equation}
  Q^2
  =
  \mathcal L_v+\delta_{\rm gauge}(\Phi+\iota_v A)
  +\delta_{SU(2)_R}(\Theta)+\delta_{\rm flavor}(m),
  \label{eq:s4-q-square-data}
\end{equation}
on gauge-covariant fields, with flavor masses included when hypermultiplets are
present.

\begin{hypothesis}[Pestun localizing cycle and square identity]
\label{hyp:s4-pestun-localizing-cycle}
The round-sphere vector-multiplet localizing functional is
\begin{equation}
  V_{\rm vec}
  =
  \int_{S^4}\sqrt g\,\operatorname{tr}
  \left(
    \Psi^\dagger Q\Psi+\eta^\dagger Q\eta+\chi^\dagger Q\chi
  \right),
  \label{eq:s4-vector-localizing-functional}
\end{equation}
with the real cycle chosen so that \(A\) is a unitary connection, the scalar
\(\Phi\) in \(Q^2\) is the real Coulomb coordinate, the remaining real scalar
direction is denoted by \(Y\), and the auxiliary fields are integrated along
the contour making the Gaussian convergent.  After eliminating \(H\), the
spinor bilinears determine nonnegative coefficient functions on \(S^4\).
The required bosonic square identity is
\begin{equation}
\begin{aligned}
  (QV_{\rm vec})_{\rm bos}
  =
  \int_{S^4}\sqrt g\,\operatorname{tr}\Big(
  &|\iota_vF+D_A\Phi|^2
  +|D_A\bar\Phi|^2
  +|[\Phi,\bar\Phi]|^2
  +|D_A Y|^2+\frac1{r^2}|Y|^2\\
  &+c_+(x)|F^+ + \mathcal T^+(x)Y|^2
  +c_-(x)|F^- + \mathcal T^-(x)Y|^2\\
  &+|D_{AB}-D_{AB}(\Phi,\bar\Phi,Y)|^2
  \Big),
\end{aligned}
  \label{eq:s4-localizing-square-decomposition}
\end{equation}
where \(c_\pm(x)\ge0\), \(c_+\) vanishes only at one pole, and \(c_-\) vanishes
only at the other pole.  In polar angle \(\theta\) from the north pole one may
normalize them as positive multiples of
\((1-\cos\theta)/2\) and \((1+\cos\theta)/2\).  The two-forms
\(\mathcal T^\pm\) are the self-dual and anti-self-dual curvature-coupling
bilinears made from the spinors in
\eqref{eq:s4-conformal-killing-spinor-system}; on the patches where the
corresponding spinor norm is nonzero they can be represented by
\(\xi^A\sigma_{\mu\nu}\xi_A\) and
\(\bar\xi^A\bar\sigma_{\mu\nu}\bar\xi_A\), multiplied by the round-sphere
curvature scale.  The weighted products in
\eqref{eq:s4-localizing-square-decomposition} are smooth at the poles.  The
last term records all auxiliary-field curvature couplings as the algebraic
square of \(D_{AB}\) around the value fixed by
\eqref{eq:s4-vector-susy-lambda}--\eqref{eq:s4-vector-susy-lambdabar}.  The
hypermultiplet localizing term is the corresponding positive sum of the
Dirac-type \(Q\)-variation squares.  Formula
\eqref{eq:s4-localizing-square-decomposition} is a field-theoretic hypothesis
inside the localization structure: without this off-shell square identity and
cycle, the Cartan matrix integral is only an ansatz.
\end{hypothesis}

\paragraph{Smooth bulk \(S^4\) fixed-locus consequence.}
\label{par:s4-smooth-bulk-fixed-locus}
Assume the localizing cycle and square identity of
Hypothesis~\ref{hyp:s4-pestun-localizing-cycle}, and restrict to smooth fields
on the trivial \(G\)-bundle.  The zero locus of \(QV_{\rm vec}\), modulo
ordinary gauge transformations, is
\begin{equation}
  A_\mu=0,\qquad
  \Phi=a\in\mathfrak t,\qquad
  Y=0,\qquad
  \bar\Phi=\bar a,\qquad
  D_{AB}=D_{AB}(a,\bar a,0),
  \label{eq:s4-bulk-fixed-locus}
\end{equation}
with quotient by the Weyl group and with hypermultiplet fields set to zero
unless the mass and Coulomb parameters leave additional zero modes.

Indeed, every square in \eqref{eq:s4-localizing-square-decomposition} must
vanish.  The
\(|Y|^2/r^2\) square gives \(Y=0\), so the shifted curvature equations reduce
to \(F^+=0\) on the open set where \(c_+>0\) and \(F^-=0\) on the open set
where \(c_->0\).  These open sets cover the complements of opposite poles, so
both chiralities of \(F\) vanish on a dense open subset of \(S^4\).  Smoothness
and the smooth weighted pole behavior in
Hypothesis~\ref{hyp:s4-pestun-localizing-cycle} extend the vanishing through
the poles, giving \(F=0\) everywhere.  On the trivial bundle a smooth flat
connection on \(S^4\) is globally gauge-equivalent to \(A=0\).  The scalar
equations give \(d\Phi=d\bar\Phi=0\) in that gauge and
\([\Phi,\bar\Phi]=0\), hence the constant scalar may be conjugated into
\(\mathfrak t\).  Reducible points are not extra saddles; they are points where
the stabilizer of \(a\) is larger than the maximal torus, and their gauge zero
modes are accounted for by the Faddeev--Popov determinant and the Weyl quotient.
The auxiliary equation is algebraic and gives \(D_{AB}(a,\bar a,0)\).  The
positive hypermultiplet squares force the hyper fields to vanish away from the
usual mass/Coulomb zero-mode walls, where the determinant must be interpreted
with the zero modes separated.

Nonzero instanton number is a different sector, not an extra smooth bulk
deformation of the smooth bulk locus \eqref{eq:s4-bulk-fixed-locus}.  If the
second Chern number is nonzero, \(F\) cannot vanish globally.  In the
large-localizing-parameter limit the positive curvature terms suppress
curvature on compact subsets away from the zeros of \(v\), while the topological
charge remains fixed.  Uhlenbeck compactness then leaves only pointlike
curvature concentration at the north and south poles.  After rescaling a small
ball around a pole to its tangent \(\mathbb C^2\), the equations reduce to the
ordinary equivariant instanton equation at one pole and anti-instanton equation
at the other.  The choice of Gieseker, Uhlenbeck, or another compactification
of these tangent-space moduli problems is therefore the separate input recorded
in Definition~\ref{def:s4-nekrasov-factor}; it is not derived by the smooth bulk
fixed-locus argument.

The reduction has three logically different physics layers.  The first layer is
the off-shell \(QV\) square identity and the real cycle; it controls the smooth
bulk equations and identifies the Coulomb saddle.  The second layer is the
topological charge sector; it explains why nonzero instanton number survives as
pole-local tunneling data rather than as a smooth deformation of the Coulomb
configuration.  The third layer is the perturbative normal-mode calculation
around the Coulomb saddle.  Only this third layer produces the spectrum whose
finite part is written with \(H\).  Thus the one-loop product below is placed
after the fixed-equation analysis: its Barnes algebra regularizes a determinant
whose normal complex has already been supplied by the gauge theory.  This
ordering is part of the QFT content.  A finite product identity can verify
counterterm conventions, root-pair cancellations, and hypermultiplet signs, but
the matrix integral obtains field-theoretic meaning from the preceding
localization structure and from the continuum comparison datum in
Definition~\ref{def:compact-supersymmetric-localization-datum}.
For that reason the next lemma is intentionally modest.  It fixes the analytic
normalization of one entire function after the equivariant index has produced
the mode multiplicities.  It does not choose the integration cycle, select the
instanton compactification, or justify the continuum limit; those pieces remain
the localization data just described.  This keeps the analytic determinant
calculation in its proper physical role, as a consequence of the normal
complex rather than the source of the localization argument.

Define the even entire function
\begin{equation}
  H(x)
  =
  \prod_{n=1}^{\infty}
  \left(1+\frac{x^2}{n^2}\right)^n
  \exp\!\left(-\frac{x^2}{n}\right),
  \label{eq:pestun-H-function}
\end{equation}
where the product is taken in the displayed counterterm convention.

\begin{lemma}[Local uniform convergence and Barnes representation of \(H\)]
\label{lem:s4-H-function-convention}
The product in \eqref{eq:pestun-H-function} converges locally uniformly on
the complex \(x\)-plane, defines an even entire function, and satisfies
\begin{equation}
  \frac{d}{dx}\log H(x)
  =
  \sum_{n=1}^{\infty}
  \left(
    \frac{2nx}{n^2+x^2}-\frac{2x}{n}
  \right)
  \label{eq:H-log-derivative}
\end{equation}
away from its zeros.  If \(G_B\) is the Barnes \(G\)-function normalized by
\(G_B(1)=1\) and \(G_B(z+1)=\Gamma(z)G_B(z)\), then
\begin{equation}
  H(x)
  =
  \exp\!\bigl(-(1+\gamma_{\mathrm E})x^2\bigr)\,
  G_B(1+\ii x)G_B(1-\ii x).
  \label{eq:H-Barnes-G-convention}
\end{equation}
Changing the exponential prefactor \(\exp(c x^2)\) is a local quadratic
counterterm for the background vector multiplet or the gauge coupling; it is
therefore a convention that must be kept fixed when comparing Gaussian
coefficients.
\end{lemma}

\begin{proof}
Let \(K\subset\mathbb C\) be compact and choose \(N\) so that
\(|x|^2/n^2\leq 1/2\) for \(x\in K\) and \(n\geq N\).  For \(n\geq N\), the
logarithm of the \(n\)-th factor is
\[
  n\log\!\left(1+\frac{x^2}{n^2}\right)-\frac{x^2}{n}
  =
  -\frac{x^4}{2n^3}+R_n(x),
  \qquad
  |R_n(x)|\leq C_K n^{-5},
\]
uniformly on \(K\).  The series of logarithms therefore converges uniformly on
\(K\) after the finitely many initial factors are separated.  The resulting
Weierstrass product converges locally uniformly and defines an entire
function.  Each factor depends only on \(x^2\), so the function is even; its
zeros are at \(x=\pm \ii n\) with the multiplicities encoded by the exponent
\(n\).

On a compact set avoiding these zeros, the differentiated logarithmic series
is uniformly convergent because
\[
  \frac{d}{dx}
  \left[
    n\log\!\left(1+\frac{x^2}{n^2}\right)-\frac{x^2}{n}
  \right]
  =
  \frac{2nx}{n^2+x^2}-\frac{2x}{n}
  =
  O_K(n^{-3}) .
\]
Termwise differentiation is then justified by the Weierstrass theorem for
locally uniformly convergent holomorphic series and gives
\eqref{eq:H-log-derivative}.

For the Barnes representation, use the standard Weierstrass product
\[
  G_B(1+z)
  =
  (2\pi)^{z/2}
  \exp\!\left[-\frac12\bigl(z+(1+\gamma_{\rm E})z^2\bigr)\right]
  \prod_{n=1}^\infty
  \left(1+\frac zn\right)^n
  \exp\!\left(\frac{z^2}{2n}-z\right),
\]
which fixes the normalization \(G_B(1)=1\) and the functional equation
\(G_B(z+1)=\Gamma(z)G_B(z)\).  Multiplying the products for \(z=\ii x\) and
\(z=-\ii x\), the factors \((2\pi)^{\pm \ii x/2}\) and \(\exp(\mp \ii x/2)\)
cancel, while
\[
  \left(1+\frac{\ii x}{n}\right)^n
  \left(1-\frac{\ii x}{n}\right)^n
  \exp\!\left(-\frac{x^2}{n}\right)
  =
  \left(1+\frac{x^2}{n^2}\right)^n
  \exp\!\left(-\frac{x^2}{n}\right).
\]
The remaining exponential is \(\exp((1+\gamma_{\rm E})x^2)\), which gives
\eqref{eq:H-Barnes-G-convention}.  Multiplying \(H\) by
\(\exp(c x^2)\) changes only this local quadratic prefactor.
\end{proof}

The one-loop determinant formula below comes from the gauge-fixed normal
complex around the Coulomb locus, not from a free choice of exponents.  At
\(A=0,\Phi=a\), decompose every field into Cartan directions and
root/weight sectors.  Cartan scalar zero modes are tangent to the fixed locus;
Cartan gauge rotations are divided by the gauge volume.  For an ordered
nonzero root \(\alpha\), the vector-multiplet normal complex has principal
part
\begin{equation}
  \Omega^0_\alpha(S^4)
  \xrightarrow{\,d\,}
  \Omega^1_\alpha(S^4)
  \xrightarrow{\,P_v^+d\,}
  E^+_{v,\alpha}(S^4),
  \label{eq:s4-vector-normal-symbol-complex}
\end{equation}
where the first arrow is the infinitesimal gauge action, \(E_v^+\) is the
self-dual two-form bundle rotated by the chosen spinor bilinears, and the
lower-order terms are the \(Q^2\)-weights
\(\mathcal L_v+\ii\alpha(a)+\Theta\).  For a hypermultiplet weight \(\rho\),
the normal complex is the corresponding \(Q\)-Dirac complex with equivariant
mass \(\ii\rho(a)+\ii m\).  These complexes are transversally elliptic for the
\(U(1)^2\) action generated by \(v\): for every nonzero covector \(u\) normal
to the orbit direction, the symbol sequence of
\eqref{eq:s4-vector-normal-symbol-complex} is exact.  In a frame
\(u=e^1\), this is the elementary exact sequence
\[
  \mathbb C
  \xrightarrow{\lambda\mapsto\lambda e^1}
  \mathbb C^4
  \xrightarrow{a\mapsto P^+(e^1\wedge a)}
  \mathbb C^3,
\]
whose second kernel is precisely the first image.

The equivariant index of this normal complex gives the residual unpaired
cohomology after acyclic \(Q\)-doublets have cancelled.  In the round
specialization \(\epsilon_1=\epsilon_2=1/r\), the one-variable character of one
ordered vector root sector and one full hypermultiplet weight sector is
\begin{equation}
\begin{aligned}
  \chi_{\alpha}^{\rm vec}(q)
  &=
  \frac{q}{(1-q)^2}\,e^{\ii r\alpha(a)}
  =
  \sum_{n\ge1}n\,q^n e^{\ii r\alpha(a)},\\
  \chi_{\rho}^{\rm hyp}(q)
  &=
  -\frac{q}{(1-q)^2}\,e^{\ii(r\rho(a)+rm)}
  =
  -\sum_{n\ge1}n\,q^n e^{\ii(r\rho(a)+rm)} .
\end{aligned}
  \label{eq:s4-normal-complex-index-multiplicities}
\end{equation}
The sign distinguishes numerator and denominator contributions to the
superdeterminant.  Combining the ordered roots \(\alpha\) and \(-\alpha\)
turns the two linear factors at each level into \(n^2+r^2\alpha(a)^2\) with
multiplicity \(2n\) for each positive root.  Thus the mode multiplicity \(n\)
is the coefficient of the transversally elliptic index, while the finite
product below is only the finite-part regularization of the determinant after
that index has been supplied.

The finite-part operation is as follows.  After pairing all acyclic
\(Q\)-doublets in a fixed root or weight sector, the unpaired finite-cutoff
determinant has the form
\begin{equation}
  D_{\kappa,N}^{\rm raw}(x)
  =
  \prod_{n=1}^{N}(n^2+x^2)^{\kappa n},
  \qquad \kappa\in\mathbb Z .
  \label{eq:s4-raw-mode-determinant-product}
\end{equation}
Here \(x\) is the dimensionless equivariant mass in that sector:
\(x=r\,\alpha(a)\) for a vector root, or \(x=r\rho(a)+rm\) for a hypermultiplet
weight.  The exponent \(n\) is the residual cohomological multiplicity in
\eqref{eq:s4-normal-complex-index-multiplicities}.  The constant factor
\(\prod_{n=1}^{N}n^{2\kappa n}\) is independent of the Coulomb variable and
is removed by the normalization of the partition function.  The remaining
Coulomb-dependent finite part is
\begin{equation}
  D_{\kappa,N}^{\rm fp}(x)
  =
  \prod_{n=1}^{N}
  \left(1+\frac{x^2}{n^2}\right)^{\kappa n}
  \exp\!\left(-\kappa\frac{x^2}{n}\right)
  =
  H_N(x)^{\kappa},
  \label{eq:s4-finite-part-mode-determinant}
\end{equation}
where \(H_N\) is the \(N\)-th partial product of
\eqref{eq:pestun-H-function}.  The exponential in
\eqref{eq:s4-finite-part-mode-determinant} subtracts the logarithmically
divergent local term
\(\kappa x^2\sum_{n\le N}n^{-1}\).  Since \(x\) is linear in the background
scalar and mass parameters, this subtraction is a quadratic local
counterterm in the rigid vector or flavor background.  The finite determinant
is therefore the finite part specified by the quadratic counterterm convention in
Lemma~\ref{lem:s4-H-function-convention}.

With this convention the perturbative \(S^4\) determinant is obtained from a
short sector calculation.  A vector root pair \(\{\alpha,-\alpha\}\) contributes
the Cartan gauge-fixing factor \(\alpha(a)^2\), the orientation of the real
determinant line fixed by pairing conjugate root sectors, and the finite part
with \(\kappa=2\), hence \(H(r\alpha(a))^2\) for each positive root.  A full
hypermultiplet weight \(\rho\) contributes the inverse finite part with
\(\kappa=-1\), hence \(H(r\rho(a)+rm)^{-1}\).  Zero modes at walls such as
\(\alpha(a)=0\) are not included in the determinant; they are separated before
the Weyl quotient and the Cartan contour are defined.  Remaining determinant
phases and shifts by \(\exp(c x^2)\) are local gauge/flavor/gravitational
counterterm choices.  The formula below is this normal-complex sector
calculation multiplied over the root and weight multisets.  The continuum QFT
still has to provide the regulated field space, the \(Q\)-Stokes theorem, and
the absence or control of boundary contributions in the chosen integration
cycle; those structures are part of
Definition~\ref{def:compact-supersymmetric-localization-datum}.

For a vector multiplet and hypermultiplets of masses \(m_\ell\) in
representations \(R_\ell\), the perturbative determinant on the round sphere is
\begin{equation}
  Z_{1{\rm -loop}}^{S^4}(a)
  =
  \prod_{\alpha\in\Delta_+}
  \alpha(a)^2 H(r\,\alpha(a))^2
  \prod_\ell\prod_{\rho\in{\rm wt}(R_\ell)}
  H\!\left(r\,\rho(a)+r\,m_\ell\right)^{-1}.
  \label{eq:s4-localization-one-loop}
\end{equation}
The factor \(\prod_{\alpha>0}\alpha(a)^2\) is the Cartan gauge-fixing
Jacobian.  If the representation is real or pseudoreal, the weights are
counted with the multiplicity appropriate to a full hypermultiplet in the
chosen convention.  A different convention for the hypermultiplet mass shifts
the argument of \(H\) by the corresponding \(R\)-symmetry background; this is
a coordinate change in the localization structure, not a change of the underlying
determinant complex.

\paragraph{\(\mathcal N=4\) determinant cancellation on \(S^4\).}
Let \(G\) be compact connected with complexified Lie algebra
\(\mathfrak g_{\mathbb C}\), Cartan subalgebra \(\mathfrak t_{\mathbb C}\),
root system \(\Delta\), chosen positive roots \(\Delta_+\), and invariant
inner product fixed by the Yang--Mills action
\(\frac1{4g_{\rm YM}^2}\int \operatorname{tr}(F_{\mu\nu}F^{\mu\nu})\).
Consider the \(\mathcal N=2\) localization formula
\eqref{eq:s4-localization-one-loop} for one vector multiplet and one
massless adjoint hypermultiplet, with the hypermultiplet mass coordinate
tuned to the point where the field content is an \(\mathcal N=4\) vector
multiplet.  Then the non-polynomial \(H\)-factors cancel identically and the
one-loop factor is only the Cartan gauge-fixing Jacobian:
\begin{equation}
  Z_{1{\rm -loop}}^{\mathcal N=4}(a)
  =
  \prod_{\alpha\in\Delta_+}\alpha(a)^2 .
  \label{eq:s4-n4-one-loop-cancellation}
\end{equation}
The adjoint representation has weight multiset
\[
  {\rm wt}(\mathfrak g_{\mathbb C})
  =
  \Delta\;\sqcup\;\{0,\ldots,0\},
\]
where the zero weight occurs \(\operatorname{rank}G\) times.  By
Lemma~\ref{lem:s4-H-function-convention}, \(H\) is an even entire function
and \(H(0)=1\).  Therefore the zero weights of the adjoint hypermultiplet
contribute the constant \(H(0)^{-\operatorname{rank}G}=1\), while each pair
\(\{\alpha,-\alpha\}\) contributes
\[
  H(r\,\alpha(a))^{-1}H(-r\,\alpha(a))^{-1}
  =
  H(r\,\alpha(a))^{-2}.
\]
The vector multiplet contributes
\[
  \prod_{\alpha\in\Delta_+}\alpha(a)^2H(r\,\alpha(a))^2 .
\]
Multiplying the vector and adjoint-hyper factors root pair by root pair
leaves exactly \eqref{eq:s4-n4-one-loop-cancellation}.  No regularization
ambiguity is hidden in this cancellation: changing the normalization of
\(H\) by an even quadratic exponential would multiply the vector and adjoint
hypermultiplet factors by inverse local quadratic counterterms at the
\(\mathcal N=4\) point, so the cancellation holds in the determinant
convention fixed in \eqref{eq:H-Barnes-G-convention}.

\begin{definition}[Nekrasov factor in the \(S^4\) formula]
\label{def:s4-nekrasov-factor}
For the four-dimensional localization formula, \(Z_{\rm inst}(a,\tau;
\epsilon_1,\epsilon_2)\) denotes the equivariant integral over the framed
instanton moduli space on \(\mathbb C^2\), with equivariant rotation weights
\((\epsilon_1,\epsilon_2)\), Coulomb parameter \(a\), and
\[
  \tau=\frac{\theta}{2\pi}+\frac{4\pi\ii}{g_{\rm YM}^2}.
\]
For \(U(N)\) this can be defined by integration over the smooth framed
torsion-free-sheaf compactification.  For other gauge groups and matter
systems the formula requires the corresponding compactification or an
equivalent contour prescription as part of the stated localization structure.
\end{definition}

\section{Singular Instantons And Compactification Choices}
\label{sec:susy-localization-singular-instantons}

The factor \(Z_{\rm inst}\) in the \(S^4\) formula is a compactification
choice, not a decorative notation.  Smooth instantons on \(\mathbb C^2\) can
shrink to zero size.  The limiting curvature is then no longer represented
by a smooth connection; the lost action is recorded as a point with
multiplicity.  A localization formula that sums over point instantons must
therefore say which compactified moduli space is being integrated over.

\begin{definition}[Framed instanton compactifications on \(\mathbb C^2\)]
\label{def:framed-instanton-compactifications-c2}
Let \(\mathcal M_{k,N}^{\rm sm}\) be the framed smooth \(U(N)\)
\(k\)-instanton moduli space on \(\mathbb C^2\).
\begin{enumerate}
  \item The Uhlenbeck compactification is stratified by
  \begin{equation}
    \overline{\mathcal M}_{k,N}^{\,U}
    =
    \bigsqcup_{\ell=0}^{k}
    \mathcal M_{k-\ell,N}^{\rm sm}\times
    {\rm Sym}^{\ell}(\mathbb C^2).
    \label{eq:uhlenbeck-stratification-c2}
  \end{equation}
  The integer \(\ell\) is the amount of instanton number carried by ideal
  point instantons.
  \item The Gieseker, or framed torsion-free-sheaf, compactification
  \(\mathfrak M_{k,N}^{G}\) is the smooth stable ADHM quotient.  It carries
  a proper morphism
  \[
    \pi_{G\to U}:\mathfrak M_{k,N}^{G}
    \longrightarrow
    \overline{\mathcal M}_{k,N}^{\,U}
  \]
  that resolves the small-instanton strata relevant to equivariant
  localization.
  \item On a compact four-manifold \(X\), the Donaldson--Uhlenbeck
  compactification of ASD connections is the analogous compactification by
  weak limits of connections away from finitely many points, together with
  the lost second Chern class recorded as an effective zero-cycle on \(X\).
\end{enumerate}
These spaces have different meanings.  A field-theoretic path integral with
a specified regulator must explain which compactification, or which
equivalent contour, it produces.
\end{definition}

\begin{construction}[Gieseker-to-Uhlenbeck charge bookkeeping]
\label{cons:gieseker-uhlenbeck-charge-bookkeeping}
Work on the framed \(U(N)\) instanton compactification on \(\mathbb C^2\),
equivalently on \(\mathbb P^2\) with a fixed framing along the line at
infinity, and take \(c_1=0\).  Let \(E\) be a framed torsion-free sheaf in
the Gieseker compactification.  Its double dual \(E^{\ast\ast}\) is locally
free, and there is an exact sequence
\begin{equation}
  0\longrightarrow E\longrightarrow E^{\ast\ast}
  \longrightarrow Q\longrightarrow0
  \label{eq:torsion-free-double-dual-quotient}
\end{equation}
with \(Q\) a zero-dimensional sheaf of length \(\ell\).  If the instanton
number is \(k(F)=-\int_{\mathbb P^2}{\rm ch}_2(F)\) in the framed
\(c_1=0\) convention, then
\begin{equation}
  k(E)=k(E^{\ast\ast})+\ell .
  \label{eq:gieseker-uhlenbeck-charge-split}
\end{equation}
Thus the Uhlenbeck morphism sends \(E\) to the smooth framed instanton
represented by \(E^{\ast\ast}\), together with the effective length-\(\ell\)
zero-cycle supporting \(Q\).
\end{construction}

The charge identity in
\eqref{eq:gieseker-uhlenbeck-charge-split} is a finite Chern-character
calculation.  On a smooth complex surface the double dual of a torsion-free
sheaf is locally free, and the quotient in
\eqref{eq:torsion-free-double-dual-quotient} is supported in complex
codimension two.  The Chern character is additive in exact sequences, so
\[
  {\rm ch}(E^{\ast\ast})={\rm ch}(E)+{\rm ch}(Q).
\]
Since \(Q\) is zero-dimensional of length \(\ell\), its only nonzero Chern
character component is \({\rm ch}_2(Q)=\ell[{\rm pt}]\).  Therefore
\[
  {\rm ch}_2(E)={\rm ch}_2(E^{\ast\ast})-\ell[{\rm pt}].
\]
Integrating and multiplying by the instanton-number sign convention gives
\[
  k(E)
  =
  -\int{\rm ch}_2(E)
  =
  -\int{\rm ch}_2(E^{\ast\ast})+\ell
  =
  k(E^{\ast\ast})+\ell .
\]
The support cycle of \(Q\) is an element of
\({\rm Sym}^{\ell}(\mathbb C^2)\).  Keeping only
\((E^{\ast\ast},{\rm supp}\,Q)\) gives the corresponding Uhlenbeck stratum,
while the Gieseker space retains the punctual quotient data that resolve that
stratum.

\begin{remark}[Dimension of Uhlenbeck strata]
\label{rem:uhlenbeck-stratum-dimension}
The complex dimension of the top stratum
\(\mathcal M_{k,N}^{\rm sm}\) is \(2kN\).  The stratum with \(\ell\) ideal
points in \eqref{eq:uhlenbeck-stratification-c2} has complex dimension
\[
  2(k-\ell)N+2\ell .
\]
Hence its complex codimension inside the top-dimensional instanton
component is
\begin{equation}
  2\ell(N-1).
  \label{eq:uhlenbeck-stratum-codimension}
\end{equation}
For \(N=2\) and \(\ell=1\), this is real codimension four.

The ADHM dimension count in
Chapter~\ref{ch:susy-four-dimensional-n2-seiberg-witten} gives complex dimension
\(2rN\) for the smooth framed charge-\(r\) moduli space.  The symmetric
product \({\rm Sym}^{\ell}(\mathbb C^2)\) has complex dimension \(2\ell\).
Adding the two dimensions gives the displayed stratum dimension with
\(r=k-\ell\).  Subtracting from \(2kN\) gives
\[
  2kN-\bigl(2(k-\ell)N+2\ell\bigr)=2\ell(N-1).
\]
\end{remark}

\begin{definition}[ADHM stability and the small-instanton locus]
\label{def:adhm-stability-small-instanton}
Let \(K\simeq\mathbb C^k\), \(W\simeq\mathbb C^N\), and let
\((B_1,B_2,I,J)\) obey the complex ADHM equation
\[
  [B_1,B_2]+IJ=0 .
\]
The ADHM tuple is stable if no proper subspace
\(S\subsetneq K\) satisfies
\[
  B_1S\subset S,\qquad B_2S\subset S,\qquad I(W)\subset S .
\]
The Gieseker compactification is the stable GIT quotient by \(GL(K)\), or
equivalently the positive real-moment-map hyperkähler quotient.  The
Uhlenbeck compactification is obtained from the affine quotient after
discarding the stable framing data carried by punctual quotients.  Failure
of stability is the ADHM description of the small-instanton stratum.
\end{definition}

The geometric content of this definition is as follows.  If the cyclic span of
\(I(W)\) under \(B_1\) and \(B_2\) is all of \(K\), the
framing data generate the length-\(k\) torsion-free sheaf and the
complexified gauge action has the expected stable orbit.  If the cyclic span
is a proper subspace \(S\), then the quotient \(K/S\) carries ADHM length not
seen by the framed smooth connection.  In the affine quotient this quotient
is recorded only by the eigenvalue data of \(B_1,B_2\), hence by points of
\(\mathbb C^2\) with multiplicity.  These are precisely the ideal
point-instanton coordinates in the Uhlenbeck stratum.  The positive
real-moment-map quotient forbids this loss by forcing stability; the price is
that the singular stratum is replaced by the Gieseker resolution.

\begin{proposition}[Positive ADHM moment map excludes the small-instanton stratum]
\label{prop:positive-adhm-moment-map-stability}
Let \(K\) and \(W\) be Hermitian vector spaces and let
\((B_1,B_2,I,J)\in{\rm End}(K)^{\oplus2}\oplus{\rm Hom}(W,K)
\oplus{\rm Hom}(K,W)\) obey the ADHM equations
\begin{equation}
  [B_1,B_2]+IJ=0,
  \label{eq:positive-adhm-complex-moment-map}
\end{equation}
\begin{equation}
  [B_1,B_1^\dagger]+[B_2,B_2^\dagger]
  +II^\dagger-J^\dagger J
  =
  \zeta\,{\bf 1}_K ,
  \qquad \zeta>0 .
  \label{eq:adhm-real-moment-map-positive}
\end{equation}
Then the ADHM configuration is stable in the sense of
Definition~\ref{def:adhm-stability-small-instanton}.  Equivalently, the
positive real moment-map parameter forbids a quotient \(K/S\) that is
invisible to the framing \(I(W)\), hence forbids the finite-dimensional
small-instanton degeneration represented by failure of stability.
\end{proposition}

\begin{proof}
Suppose, for contradiction, that there is a proper subspace
\(S\subsetneq K\) such that \(B_iS\subset S\) for \(i=1,2\) and
\(I(W)\subset S\).  Let \(T=S^\perp\), and write each \(B_i\) in block form
with respect to \(K=S\oplus T\):
\[
  B_i=
  \begin{pmatrix}
    A_i&C_i\\
    0&D_i
  \end{pmatrix}.
\]
The \(T\)-trace of the commutator block is
\begin{equation}
  \operatorname{Tr}_T\!\bigl([B_i,B_i^\dagger]_{TT}\bigr)
  =
  \operatorname{Tr}(D_iD_i^\dagger)
  -\operatorname{Tr}(C_i^\dagger C_i+D_i^\dagger D_i)
  =
  -\operatorname{Tr}(C_i^\dagger C_i).
  \label{eq:adhm-stability-tt-commutator-trace}
\end{equation}
Since \(I(W)\subset S\), the \(T\)-block of \(II^\dagger\) is zero.  The
\(T\)-trace of \(-J^\dagger J\) is
\(-\operatorname{Tr}_T(J^\dagger J)\le0\).  Taking the trace over \(T\) in
\eqref{eq:adhm-real-moment-map-positive} therefore gives
\begin{equation}
  -\sum_{i=1}^2\operatorname{Tr}(C_i^\dagger C_i)
  -\operatorname{Tr}_T(J^\dagger J)
  =
  \zeta\,\dim T .
  \label{eq:adhm-positive-fi-contradiction}
\end{equation}
The left hand side is nonpositive and the right hand side is positive
because \(\zeta>0\) and \(T\ne0\).  This contradiction proves stability.
\end{proof}

\begin{remark}[What this finite-dimensional statement does not prove]
\label{rem:positive-adhm-does-not-prove-field-theory-matching}
Proposition~\ref{prop:positive-adhm-moment-map-stability} is a theorem
inside the ADHM finite-dimensional reduction.  It justifies the common
statement that a positive ADHM/FI resolution, for example the
finite-dimensional resolution induced by noncommutative deformation in the
Nekrasov--Schwarz construction, removes the small-instanton stratum.  It
does not by itself prove that a four-dimensional supersymmetric gauge-theory
path integral with a given regulator localizes to this resolved quotient.
That identification is the separate regulator theorem, or hypothesis, stated
in Hypothesis~\ref{hyp:field-theoretic-gieseker-matching}.
\end{remark}

\begin{definition}[Tangent Euler class at a Young-diagram fixed point]
\label{def:nekrasov-tangent-euler-class}
Let a fixed point of \(\mathfrak M_{k,N}^{G}\) be labeled by
\(\vec Y=(Y_1,\ldots,Y_N)\).  For a box \(s=(i,j)\in Y_\alpha\), define the
arm and relative leg lengths by
\[
  A_{Y_\alpha}(s)=Y_{\alpha,i}-j,
  \qquad
  L_{Y_\beta}(s)=Y_{\beta,j}^{\,T}-i .
\]
With Coulomb weights \(a_\alpha\) and rotation weights
\(\epsilon_1,\epsilon_2\), use the orientation convention
\begin{equation}
\begin{aligned}
  e_T(T_{\vec Y}\mathfrak M_{k,N}^{G})
  &=
  \prod_{\alpha,\beta=1}^{N}
  \prod_{s\in Y_\alpha}
  \Bigl(a_\alpha-a_\beta
  -\epsilon_1L_{Y_\beta}(s)
  +\epsilon_2(A_{Y_\alpha}(s)+1)\Bigr)
\\
  &\quad\times
  \Bigl(a_\alpha-a_\beta
  +\epsilon_1(L_{Y_\beta}(s)+1)
  -\epsilon_2A_{Y_\alpha}(s)\Bigr).
  \label{eq:nekrasov-tangent-euler-class}
\end{aligned}
\end{equation}
The pure vector fixed-point contribution is
\[
  Z_{\vec Y}^{\rm vec}
  =
  e_T(T_{\vec Y}\mathfrak M_{k,N}^{G})^{-1}
\]
in this convention.  Matter multiplets multiply this by the Euler classes of
their corresponding equivariant index bundles.
\end{definition}

\begin{proposition}[Rank-one monomial-ideal tangent character]
\label{prop:rank-one-hilbert-tangent-character}
Let \(Y\) be a Young diagram and let
\(I_Y\subset R=\mathbb C[z_1,z_2]\) be the monomial ideal whose quotient has
basis
\[
  Q_Y=R/I_Y,\qquad
  \{z_1^{i-1}z_2^{j-1}:(i,j)\in Y\}.
\]
Let \(z_1\) and \(z_2\) have torus weights
\(\epsilon_1\) and \(\epsilon_2\), and write \(t_a=e^{\epsilon_a}\).  At the
fixed point \(I_Y\in{\rm Hilb}^{|Y|}(\mathbb C^2)\), the rank-one tangent
character is
\begin{equation}
  \operatorname{ch}_{T}T_{I_Y}{\rm Hilb}^{|Y|}(\mathbb C^2)
  =
  \sum_{s\in Y}
  \left(
    t_1^{-L_Y(s)}t_2^{A_Y(s)+1}
    +
    t_1^{L_Y(s)+1}t_2^{-A_Y(s)}
  \right).
  \label{eq:rank-one-hilbert-tangent-character}
\end{equation}
Equivalently, the additive tangent weights are
\[
  -L_Y(s)\epsilon_1+(A_Y(s)+1)\epsilon_2,
  \qquad
  (L_Y(s)+1)\epsilon_1-A_Y(s)\epsilon_2,
  \qquad s\in Y .
\]
Their product is the \(N=1\) specialization of
\eqref{eq:nekrasov-tangent-euler-class}.
\end{proposition}

\begin{proof}
Set
\[
  V_Y=\operatorname{ch}_{T}Q_Y
  =
  \sum_{(i,j)\in Y}t_1^{i-1}t_2^{j-1},
  \qquad
  V_Y^\vee=\sum_{(i,j)\in Y}t_1^{1-i}t_2^{1-j}.
\]
The monomial-ideal deformation complex, equivalently the rank-one ADHM
deformation complex with framing character \(1\), gives the finite
equivariant \(K\)-theory identity
\begin{equation}
  \operatorname{ch}_{T}T_{I_Y}{\rm Hilb}^{|Y|}(\mathbb C^2)
  =
  V_Y^\vee+t_1t_2V_Y-(1-t_1)(1-t_2)V_YV_Y^\vee .
  \label{eq:rank-one-hilbert-k-character}
\end{equation}
Here the last term subtracts the gauge and relation part of the two-arrow
ADHM complex, while \(V_Y^\vee\) and \(t_1t_2V_Y\) are the two framing
terms.  Expanding \((1-t_1)(1-t_2)V_YV_Y^\vee\) is a finite second-difference
calculation on the set of box displacements in \(Y-Y\): every displacement
whose four adjacent translates stay inside the diagram cancels, and the only
uncancelled displacements are the two hooks that cross the boundary at each
box \(s\).  The east hook contributes
\(t_1^{-L_Y(s)}t_2^{A_Y(s)+1}\), and the south hook contributes
\(t_1^{L_Y(s)+1}t_2^{-A_Y(s)}\).  Summing over boxes gives
\eqref{eq:rank-one-hilbert-tangent-character}.  Passing from multiplicative
characters to additive weights gives the displayed two weights per box, and
their product is exactly \eqref{eq:nekrasov-tangent-euler-class} with
\(N=1\) and \(a_\alpha-a_\beta=0\).
\end{proof}

For \(k=1\), the fixed points are the \(N\) choices in which one Young
diagram \(Y_\alpha\) consists of a single box.  Formula
\eqref{eq:nekrasov-tangent-euler-class} gives
\[
  Z_1^{U(N)}
  =
  \frac1{\epsilon_1\epsilon_2}
  \sum_{\alpha=1}^{N}
  \prod_{\beta\ne\alpha}
  \frac1{(a_\alpha-a_\beta)
  (a_\alpha-a_\beta+\epsilon_1+\epsilon_2)} ,
\]
which is the one-instanton formula used in
Definition~\ref{def:pure-vector-nekrasov-partition}.
Let \(s=(1,1)\) be the single box of \(Y_\alpha\).  For \(\beta=\alpha\),
\[
  A_{Y_\alpha}(s)=0,\qquad L_{Y_\alpha}(s)=0,
\]
so the two tangent weights are \(\epsilon_2\) and \(\epsilon_1\).  For
\(\beta\ne\alpha\), the diagram \(Y_\beta\) is empty, hence
\[
  A_{Y_\alpha}(s)=0,\qquad L_{Y_\beta}(s)=-1.
\]
The two weights in \eqref{eq:nekrasov-tangent-euler-class} are then
\[
  a_\alpha-a_\beta+\epsilon_1+\epsilon_2,
  \qquad
  a_\alpha-a_\beta .
\]
Multiplying the weights and summing over the \(N\) possible positions of the
one box gives the displayed expression.

\begin{example}[The two-point Gieseker fixed-point laboratory]
\label{ex:k-two-hilbert-scheme-fixed-points}
The first place where the resolved punctual data are not just a projective
orientation space is the rank-one length-two Gieseker sector
\[
  \mathfrak M_{2,1}^{G}\simeq {\rm Hilb}^{2}(\mathbb C^2).
\]
This is a finite-dimensional model of the resolved ideal-point sector, not a
claim that ordinary commutative \(U(1)\) Yang--Mills has smooth finite-action
instantons.  The Uhlenbeck record over a double point at the origin sees only
the support cycle \(2[0]\in{\rm Sym}^{2}(\mathbb C^2)\).  The Gieseker space
retains the punctual quotient data; over the double point the exceptional
fiber is \(\mathbb P^1\).  The torus \(T^2_{\epsilon_1,\epsilon_2}\) has two
fixed points, corresponding to the partitions \((2)\) and \((1,1)\).
Proposition~\ref{prop:rank-one-hilbert-tangent-character} computes their
tangent characters directly from the monomial ideals.

For the horizontal partition \(Y=(2)\), the boxes have
\[
  (A,L)=((1,0),(0,0)),
\]
and \eqref{eq:nekrasov-tangent-euler-class} gives tangent weights
\[
  2\epsilon_2,\quad \epsilon_1-\epsilon_2,\quad
  \epsilon_2,\quad \epsilon_1 .
\]
Hence
\[
  e_T(T_{(2)}{\rm Hilb}^{2}(\mathbb C^2))
  =
  2\epsilon_1\epsilon_2^2(\epsilon_1-\epsilon_2).
\]
For the vertical partition \(Y=(1,1)\), the boxes have
\[
  (A,L)=((0,1),(0,0)),
\]
and the weights are
\[
  \epsilon_2-\epsilon_1,\quad 2\epsilon_1,\quad
  \epsilon_2,\quad \epsilon_1 ,
\]
so
\[
  e_T(T_{(1,1)}{\rm Hilb}^{2}(\mathbb C^2))
  =
  2\epsilon_1^2\epsilon_2(\epsilon_2-\epsilon_1).
\]
The equivariant integral of \(1\) over the resolved length-two sector is
therefore
\begin{equation}
\begin{aligned}
  \int_{{\rm Hilb}^{2}(\mathbb C^2)}1
  &=
  \frac{1}{2\epsilon_1\epsilon_2^2(\epsilon_1-\epsilon_2)}
  +
  \frac{1}{2\epsilon_1^2\epsilon_2(\epsilon_2-\epsilon_1)}
\\
  &=
  \frac{1}{2\epsilon_1^2\epsilon_2^2}.
  \label{eq:hilb-two-equivariant-integral}
\end{aligned}
\end{equation}
This calculation displays the singular-instanton bookkeeping in its smallest
nontrivial form.  The Uhlenbeck compactification remembers the effective
cycle \(2[0]\); the Gieseker/Nekrasov resolution remembers which punctual
monomial ideal lies over that cycle.  A localization formula that uses the
two Young-diagram residues is therefore using resolved punctual data, and a
field-theoretic regulator must still explain why this resolved sector, rather
than the collapsed Uhlenbeck point, is the integration cycle it selects.
\end{example}

\begin{example}[The charge-one resolution]
\label{ex:charge-one-gieseker-resolution}
For \(k=1\), the ADHM matrices \(B_1,B_2\) are complex numbers, hence give
the center point in \(\mathbb C^2\).  The remaining data
\[
  I\in W^\ast,\qquad J\in W,\qquad IJ=0
\]
modulo the \(GL(1)\) action, with positive real moment map, give
\[
  \mathfrak M_{1,N}^{G}\simeq \mathbb C^2\times T^\ast\mathbb{CP}^{N-1}.
\]
The Uhlenbeck compactification collapses the resolved orientation/size data
over the ideal point-instanton stratum.  For \(N=2\), the collapsed stratum
has real codimension four in the top moduli component, matching
Remark~\ref{rem:uhlenbeck-stratum-dimension}.
\end{example}

\begin{construction}[Charge-one Uhlenbeck map as a nilpotent-cone resolution]
\label{cons:charge-one-uhlenbeck-nilpotent-resolution}
The charge-one example admits a completely explicit map to the Uhlenbeck
compactification.  Let \(W\simeq\mathbb C^N\).  Since \(K=\mathbb C\), the
complex ADHM equation is the scalar condition
\[
  I(J)=0,
  \qquad
  I\in W^\ast,\quad J\in W .
\]
Stability is \(I\ne0\).  Dividing by \(GL(1)=\mathbb C^\ast\) gives
\[
  (I,J)/\mathbb C^\ast
  =
  \{([I],J):[I]\in \mathbb P(W^\ast),\ I(J)=0\}
  \simeq T^\ast\mathbb P(W^\ast)
  \simeq T^\ast\mathbb{CP}^{N-1}.
\]
The Uhlenbeck map is the moment-map morphism
\begin{equation}
  \mu_{\rm Uh}:T^\ast\mathbb P(W^\ast)\longrightarrow \mathfrak{sl}(W),
  \qquad
  ([I],J)\longmapsto \nu_{I,J}:=J\otimes I,
  \label{eq:charge-one-uhlenbeck-moment-map}
\end{equation}
where \(\nu_{I,J}(w)=I(w)J\).  The equation \(I(J)=0\) gives
\[
  \operatorname{Tr}\nu_{I,J}=I(J)=0,
  \qquad
  \nu_{I,J}^2=I(J)\nu_{I,J}=0,
  \qquad
  \operatorname{rank}\nu_{I,J}\le1.
\]
Thus the image is the closure of the minimal nilpotent orbit in
\(\mathfrak{sl}(W)\).  The open subset \(J\ne0\) maps to the smooth
rank-one nilpotent orbit; the zero section \(J=0\) maps to the apex
\(\nu=0\), which is precisely the ideal point-instanton stratum in the
charge-one Uhlenbeck compactification.  Including the center
\((B_1,B_2)\in\mathbb C^2\), one has
\[
  \mathfrak M_{1,N}^{G}
  \simeq
  \mathbb C^2\times T^\ast\mathbb{CP}^{N-1}
  \longrightarrow
  \mathbb C^2\times \overline{\mathcal O}_{\rm min},
\]
and the small-instanton boundary is
\(\mathbb C^2\times\{0\}\).  This is the concrete geometric content hidden
behind the phrase ``point instanton'' in the first nontrivial instanton
sector.
\end{construction}

\begin{hypothesis}[Field-theoretic matching to the Gieseker integral]
\label{hyp:field-theoretic-gieseker-matching}
The \(S^4\) formula
\eqref{eq:pestun-s4-localization} uses the Gieseker/Nekrasov equivariant
integral at each pole.  A theorem-level derivation from the four-dimensional
gauge-theory path integral requires a regulator and integration cycle whose
localized small-instanton sector is identified with
\(\mathfrak M_{k,N}^{G}\), with the Uhlenbeck space
\(\overline{\mathcal M}_{k,N}^{\,U}\) serving as a boundary compactification
rather than as the resolved integration domain.  Noncommutative or positive
ADHM-moment-map regularization supplies a finite-dimensional resolution of
the small-instanton singularity, but the construction-level equivalence
between this resolved equivariant integral and the continuum supersymmetric
path integral is retained here as a hypothesis unless the regulator has been
specified and the limiting argument has been proven.
\end{hypothesis}

\begin{definition}[Regulator selection structure for the Gieseker sector]
\label{def:gieseker-regulator-selection-datum}
Fix \(k,N\), equivariant weights
\((\epsilon_1,\epsilon_2)\), Coulomb parameter \(a\), and the trace
normalization of this volume.  A four-dimensional supersymmetric regulator is
said to select the Gieseker instanton sector at charge \(k\) if it supplies
the following structures.
\begin{enumerate}
  \item A regulated BV localization problem
  \[
    (\mathfrak F_\delta,C_\delta,\rho_\delta,Q_\delta,V_\delta),
    \qquad \delta>0,
  \]
  satisfying the finite-regulator quantum master equation and the
  \(Q_\delta\)-Stokes condition on the integration cycle.  Here \(\delta\)
  denotes the regulator parameter, not an equivariant weight.
  \item A \(Q_\delta\)-fixed local model at the pole with coordinates
  \[
    (B_1,B_2,I,J)\in
    {\rm End}(K)^{\oplus2}\oplus{\rm Hom}(W,K)\oplus{\rm Hom}(K,W),
    \qquad \dim K=k,\quad \dim W=N,
  \]
  whose equations reduce, after a controlled finite change of coordinates,
  to
  \[
    [B_1,B_2]+IJ=0,\qquad
    [B_1,B_1^\dagger]+[B_2,B_2^\dagger]
    +II^\dagger-J^\dagger J=\zeta_\delta{\bf 1}_K+R_\delta ,
  \]
  with \(\zeta_\delta>0\) and \(R_\delta\to0\) in the norm in which the
  fixed-locus reduction is being taken.  Equivalently, the regulator may supply
  another explicit resolution cycle together with an equivariant proper map
  to \(\overline{\mathcal M}_{k,N}^{\,U}\); it must then prove that the
  resolved cycle is isomorphic to \(\mathfrak M_{k,N}^{G}\) or give the
  exact contour replacing it.
  \item An orientation and determinant-line comparison between the regulated
  normal \(Q_\delta\)-complex and the equivariant tangent/index complex on
  \(\mathfrak M_{k,N}^{G}\).  In local fixed-point coordinates this means
  that the nonzero-mode Berezinian gives the Euler class
  \eqref{eq:nekrasov-tangent-euler-class}, multiplied only by declared
  one-loop determinants and local counterterms already included in the
  localization structure.
  \item A boundary-cancellation statement for the Uhlenbeck
  small-instanton collars.  In the notation of
  Chapter~\ref{ch:tqft-bv-integration-localization}, the regulator must
  prove either
  \[
    \sum_\alpha B_{S_\alpha}(\rho_\delta)\longrightarrow0
  \]
  or an equality of this boundary functional with the difference between the
  smooth-stratum pushforward and the Gieseker-resolved pushforward.  Thus
  the resolution is part of the BV pushforward structure, not an informal
  replacement made after the path integral has already been localized.
  \item A limiting theorem for equivariantly regular test insertions
  \(\mathcal O\):
  \[
    \lim_{\delta\to0}
    \int_{C_\delta}\rho_\delta\,\mathcal O\,\ee^{-tQ_\delta V_\delta}
    =
    \int_{\mathfrak M_{k,N}^{G}}^{\rm equiv}
    \mathcal O_{\rm red}\,
    \frac{e_T(E_{\rm matter})}{e_T(T\mathfrak M_{k,N}^{G})},
  \]
  with the order of the \(t\to\infty\) localization limit and the
  \(\delta\to0\) regulator limit specified.  If the result is stated only as
  a formal power series in the instanton counting parameter, the topology of
  that formal series must be part of the statement.
\end{enumerate}
\end{definition}

This hypothesis is the point at which the monograph refuses the common
shortcut.  Equivariant localization on \(\mathfrak M_{k,N}^{G}\) is a
mathematical theorem after the resolved moduli space is chosen.  The
assertion that the original QFT path integral localizes to that particular
resolved integral is an additional field-theoretic statement.  A BV
pushforward formulation, as in
Chapter~\ref{ch:tqft-bv-integration-localization}, must include the boundary
term at the singular stratum or a regulator that replaces the stratum by a
cycle without boundary.  The singular-stratum obstruction formula
\eqref{eq:singular-stratum-localized-qme-obstruction} is the
finite-dimensional statement behind this requirement: a smooth-stratum
pushforward satisfies the residual QME only after the small-instanton boundary
functional has vanished or has been cancelled by a specified resolution,
relative cycle, or residue prescription.

Definition~\ref{def:gieseker-regulator-selection-datum} is deliberately more
structured than the phrase ``use the Nekrasov partition function''.  The
finite-dimensional ADHM statement
Proposition~\ref{prop:positive-adhm-moment-map-stability} proves the
stability part of item 2 once the fixed-locus reduction has already been
made.  It proves neither the existence of the four-dimensional regulator in
item 1, nor the determinant-line comparison in item 3, nor the
small-instanton boundary cancellation in item 4, nor the limiting statement
in item 5.  Conversely, a regulator that selected the Uhlenbeck
compactification without a compensating boundary functional would not give
the Nekrasov factor appearing in \eqref{eq:pestun-s4-localization}; it would
define a different localized problem.

\begin{openproblem}[Continuum regulator theorem for Gieseker selection]
\label{op:susy-localization-gieseker-selection-theorem}
Construct, for four-dimensional \(\mathcal N=2\) gauge theory on \(S^4\), a
BV-compatible supersymmetric regulator and integration cycle satisfying
Definition~\ref{def:gieseker-regulator-selection-datum} for every instanton
charge \(k\), or prove that the standard \(S^4\) localization formula
requires an additional nonperturbative definition whose instanton sectors are
Gieseker/Nekrasov equivariant integrals by construction.  The known
finite-dimensional ADHM resolution and the noncommutative positive-moment-map
mechanism are necessary local pieces of such a theorem, but they do not by
themselves prove the continuum field-theoretic selection statement.
\end{openproblem}

The localized partition function is
\begin{equation}
  Z_{S^4}^{\mathcal N=2}
  =
  \frac1{|W_G|}
  \int_{\mathfrak t} da\,
  Z_{\rm cl}(a)\,
  Z_{1{\rm -loop}}^{S^4}(a)\,
  Z_{\rm inst}\!\left(a,\tau;\frac1r,\frac1r\right)
  Z_{\rm inst}\!\left(a,\bar\tau;\frac1r,\frac1r\right).
  \label{eq:pestun-s4-localization}
\end{equation}
The north pole contributes instantons and the south pole contributes
anti-instantons.  The two factors are complex conjugates only when the
integration cycle and masses are chosen so that reflection positivity gives
the corresponding conjugation property.

\paragraph{Protected-insertion proof-obligation ledger.}
Equation~\eqref{eq:pestun-s4-localization} is the scalar partition-function
case of a functional statement on protected insertions.  Fix a finite
test space \(\mathcal P_N\) of \(Q\)-closed regulated insertions containing
the unit insertion and the protected Wilson-loop insertions under study.
On \(\mathcal P_N\) let
\[
  \mathcal Z_{\rm reg},\quad
  \mathcal Z_{\rm def},\quad
  \mathcal Z_{\rm normal},\quad
  \mathcal Z_{\rm res},\quad
  \mathcal Z_{\rm Nek},\quad
  \mathcal Z_{\rm loc}
  \in \mathcal P_N^\vee
\]
denote, respectively, the regulated QFT functional, the functional after
the \(Q\)-exact deformation, the functional after normal Gaussian
integration, the residual Cartan-contour functional, the north/south
ADHM--Nekrasov functional, and the final localized expression.  Define
\[
\begin{aligned}
  E_{\rm Stokes} &= \mathcal Z_{\rm reg}-\mathcal Z_{\rm def},\\
  E_{\rm normal} &= \mathcal Z_{\rm def}-\mathcal Z_{\rm normal},\\
  E_{\rm res} &= \mathcal Z_{\rm normal}-\mathcal Z_{\rm res},\\
  E_{\rm inst} &= \mathcal Z_{\rm res}-\mathcal Z_{\rm Nek},\\
  E_{\rm cont} &= \mathcal Z_{\rm Nek}-\mathcal Z_{\rm loc}.
\end{aligned}
\]
Then the comparison to the displayed localization formula is the exact
bookkeeping telescope
\begin{equation}
  \mathcal Z_{\rm reg}-\mathcal Z_{\rm loc}
  =
  E_{\rm Stokes}+E_{\rm normal}+E_{\rm res}+E_{\rm inst}+E_{\rm cont}.
  \label{eq:s4-protected-insertion-localization-budget}
\end{equation}
This identity is not, by itself, evidence for localization.  It is a ledger of
the maps that a proof must construct and estimate.  Here \(E_{\rm Stokes}\)
records finite-regulator \(Q\)-Stokes, anomaly, and
boundary defects; \(E_{\rm normal}\) records normal-complex, zero-mode, and
determinant-line defects; \(E_{\rm res}\) records the residual Cartan
contour, Weyl quotient, and fixed-locus restriction of the insertion;
\(E_{\rm inst}\) records the choice of Gieseker/Nekrasov instanton
compactification at the poles; and \(E_{\rm cont}\) records the continuum
limit and contact-term comparison.  Only after a regulator theorem gives
bounds or vanishing statements for these five entries does the ordinary
localized expression become a field-theoretic equality.  For any norm on the
finite-dimensional dual space,
\[
  \left\|\mathcal Z_{\rm reg}-\mathcal Z_{\rm loc}\right\|
  \le
  \|E_{\rm Stokes}\|+\|E_{\rm normal}\|+\|E_{\rm res}\|
  +\|E_{\rm inst}\|+\|E_{\rm cont}\|.
\]
This triangle inequality is a way of recording the remaining burden, not a
bound with content until the individual residuals have been estimated.  For
the half-BPS circular Wilson loop, the fixed-locus restriction is
\(W_{\square}|_{\rm fix}=N^{-1}\operatorname{Tr}_{\square}\ee^X\); the
Laguerre and Bessel computations below evaluate
\(\mathcal Z_{\rm loc}(W_{\square})/\mathcal Z_{\rm loc}(1)\).  They do not
by themselves prove the continuum field-theoretic vanishing of the five
residuals above.

\begin{example}[\(\mathcal N=4\) Yang--Mills degeneration]
\label{ex:s4-n4-gaussian-matrix-model}
For an \(\mathcal N=2\) vector multiplet plus an adjoint hypermultiplet with
the mass coordinate tuned to the \(\mathcal N=4\) point,
the determinant cancellation above reduces the
perturbative one-loop determinant to the Cartan gauge-fixing Jacobian.  The
remaining matrix model is
\[
  Z_{S^4}^{\mathcal N=4}
  =
  \frac1{|W_G|}
  \int_{\mathfrak t}da\,
  \exp\!\left\{
    -\frac{4\pi^2r^2}{g_{\rm YM}^2}(a,a)
  \right\}
  \prod_{\alpha\in\Delta_+}\alpha(a)^2 .
\]
For \(G=U(1)\) there is no root factor and no finite-action instanton sector
on \(\mathbb C^2\).  The partition function reduces to the ordinary Gaussian
\begin{equation}
  \int_{-\infty}^{\infty} da\,
  \exp\!\left(-\frac{4\pi^2r^2}{g_{\rm YM}^2}a^2\right)
  =
  \frac{g_{\rm YM}}{2\sqrt{\pi}\,r}.
  \label{eq:s4-u1-gaussian-localization}
\end{equation}
\end{example}

\begin{definition}[Circular Wilson-loop matrix-model coupling]
\label{def:circular-wilson-loop-matrix-coupling}
For \(G=U(N)\), let \(X\) denote the Hermitian matrix whose eigenvalues
\(x_i\) are the dimensionless arguments that appear in the localized
fundamental half-BPS circular Wilson loop,
\begin{equation}
  W_{\square}(X)
  =
  \frac1N\operatorname{Tr}_{\square}\ee^{X}
  =
  \frac1N\sum_{i=1}^N\ee^{x_i}.
  \label{eq:localized-circular-wilson-insertion}
\end{equation}
The circular-loop Gaussian coupling \(\lambda_{\rm circ}\) is defined by
writing the \(\mathcal N=4\) localized \(U(N)\) Gaussian model in the form
\begin{equation}
  \dd\mu_{\lambda_{\rm circ},N}(X)
  =
  \frac1{Z_{\lambda_{\rm circ},N}}
  \exp\!\left(
    -\frac{2N}{\lambda_{\rm circ}}\operatorname{Tr}_{\square}X^2
  \right)\dd X ,
  \label{eq:circular-wilson-gaussian-matrix-measure}
\end{equation}
or, after diagonalizing \(X\),
\begin{equation}
  \dd\mu_{\lambda_{\rm circ},N}(x)
  \propto
  \prod_{i=1}^N dx_i\,
  \exp\!\left(
    -\frac{2N}{\lambda_{\rm circ}}\sum_{i=1}^Nx_i^2
  \right)
  \prod_{i<j}(x_i-x_j)^2 .
  \label{eq:circular-wilson-eigenvalue-measure}
\end{equation}
This is a definition of the coupling that enters the exact circular-loop
answer.  Converting it to a Lagrangian symbol such as \(g_{\rm YM}^2N\)
requires using the same fundamental-matrix normalization for the scalar in
the Wilson-loop exponent and for the quadratic form in the localized
Gaussian.  Rescaling the Lie-algebra generators rescales both data; the
matrix model \eqref{eq:circular-wilson-gaussian-matrix-measure} is the
normalization-invariant statement.
\end{definition}

\paragraph{Finite-\texorpdfstring{\(N\)}{N} circular Wilson loop in the Gaussian model.}
For the \(U(N)\) Gaussian matrix model
\eqref{eq:circular-wilson-gaussian-matrix-measure}, the normalized
fundamental circular Wilson-loop insertion has the exact finite-\(N\) value
\begin{equation}
  \left\langle
    \frac1N\operatorname{Tr}_{\square}\ee^X
  \right\rangle_{\lambda_{\rm circ},N}
  =
  \frac1N
  \exp\!\left(\frac{\lambda_{\rm circ}}{8N}\right)
  L_{N-1}^{(1)}
  \!\left(-\frac{\lambda_{\rm circ}}{4N}\right),
  \label{eq:finite-n-circular-wilson-laguerre}
\end{equation}
where \(L_n^{(1)}\) is the associated Laguerre polynomial.

To derive this formula, put
\[
  \alpha=\frac{2N}{\lambda_{\rm circ}} .
\]
After diagonalizing \(X\), the one-point density is expressed by the
orthogonal-polynomial kernel for the weight \(e^{-\alpha x^2}\):
\begin{equation}
  \rho_N(x)
  =
  \frac1N e^{-\alpha x^2}
  \sum_{j=0}^{N-1}\frac{p_j(x)^2}{h_j},
  \qquad
  h_j=\int_{\mathbb R}p_j(x)^2e^{-\alpha x^2}\,dx ,
  \label{eq:finite-n-hermite-density-kernel}
\end{equation}
where \(p_j\) is any degree-\(j\) polynomial family orthogonal for this
weight.  Therefore, for an auxiliary parameter \(s\),
\begin{equation}
  W_N(s)
  =
  \left\langle
    \frac1N\operatorname{Tr}\ee^{sX}
  \right\rangle
  =
  \frac1N
  \sum_{j=0}^{N-1}
  \frac1{h_j}\int_{\mathbb R}
  e^{sx}p_j(x)^2e^{-\alpha x^2}\,dx .
  \label{eq:finite-n-wilson-kernel-sum}
\end{equation}
Choose \(p_j(x)=\alpha^{-j/2}2^{-j}H_j(\sqrt\alpha x)\), where \(H_j\) is the
physicists' Hermite polynomial.  The scale factors cancel in the ratio
appearing in \eqref{eq:finite-n-wilson-kernel-sum}.  With
\(y=\sqrt\alpha x\) and \(t=s/\sqrt\alpha\), the required normalized Hermite
integral is
\begin{equation}
  \frac{\int_{\mathbb R} e^{-y^2+ty}H_j(y)^2\,dy}
       {\int_{\mathbb R} e^{-y^2}H_j(y)^2\,dy}
  =
  \exp\!\left(\frac{t^2}{4}\right)
  L_j\!\left(-\frac{t^2}{2}\right).
  \label{eq:hermite-laguerre-integral}
\end{equation}
For completeness, recall the derivation.  The Hermite generating function is
\[
  \exp(2uy-u^2)=\sum_{j\ge0}H_j(y)\frac{u^j}{j!}.
\]
Multiplying two generating functions and integrating against
\(e^{-y^2+ty}\) gives
\[
  \sqrt\pi\,
  \exp\!\left(\frac{t^2}{4}\right)
  \exp\!\left(2uv+tu+tv\right).
\]
Taking the coefficient of \(u^jv^j/(j!)^2\) and dividing by the same
coefficient at \(t=0\), namely \(2^j j!\), gives
\eqref{eq:hermite-laguerre-integral}; the remaining finite coefficient sum is
the standard Laguerre expansion
\[
  L_j(z)=\sum_{m=0}^{j}\binom{j}{m}\frac{(-z)^m}{m!}.
  \]
Substitution into \eqref{eq:finite-n-wilson-kernel-sum} yields
\begin{equation}
  W_N(s)
  =
  \frac1N
  \exp\!\left(\frac{s^2}{4\alpha}\right)
  \sum_{j=0}^{N-1}
  L_j\!\left(-\frac{s^2}{2\alpha}\right).
  \label{eq:finite-n-wilson-laguerre-sum}
\end{equation}
The elementary identity
\[
  \sum_{j=0}^{N-1}L_j(z)=L_{N-1}^{(1)}(z)
\]
follows by summing the coefficient formula for \(L_j\) and using
\(\sum_{j=m}^{N-1}\binom jm=\binom N{m+1}\), which is precisely the
coefficient formula for \(L_{N-1}^{(1)}\).  Setting \(s=1\) and
\(\alpha=2N/\lambda_{\rm circ}\) gives
\eqref{eq:finite-n-circular-wilson-laguerre}.

\begin{quotedtheorem}[Gaussian log-gas convergence and exponential insertion]
\label{qt:gaussian-log-gas-exponential-insertion}
Use the standard Gaussian \(\beta=2\) log-gas large-deviation and edge-tail
theory, in the form recorded for example in
\cite{BenArousGuionnet1997,AndersonGuionnetZeitouni2010}.
Fix \(\lambda>0\), and let
\[
  L_N=\frac1N\sum_{i=1}^N\delta_{x_i}
\]
be the empirical measure of the eigenvalue density
\eqref{eq:circular-wilson-eigenvalue-measure}.  The sequence \(L_N\) obeys a
large-deviation principle on probability measures on \(\mathbb R\), at speed
\(N^2\), with good rate function
\[
  \mathcal I_\lambda(\nu)-\inf_\eta\mathcal I_\lambda(\eta),
  \qquad
  \mathcal I_\lambda(\nu)
  =
  \frac{2}{\lambda}\int x^2\,d\nu(x)
  -
  \iint\log|x-y|\,d\nu(x)d\nu(y),
\]
on the domain where the quadratic moment and logarithmic energy are finite,
and \(+\infty\) otherwise.  The minimizer \(\mu_\lambda\) is unique.
Consequently \(L_N\to\mu_\lambda\) in probability, and almost surely if the
finite-\(N\) ensembles are realized independently on a product probability
space.  For every bounded continuous function \(f\),
\[
  \lim_{N\to\infty}
  \mathbb E_{\lambda,N}\!\left[\int f\,dL_N\right]
  =
  \int f\,d\mu_\lambda .
\]
For the Gaussian potential the right edge has an exponential tail strong
enough to control the circular Wilson insertion:
\begin{equation}
  \lim_{R\to\infty}\limsup_{N\to\infty}
  \mathbb E_{\lambda,N}\!\left[
    \int_{|x|>R}\ee^x\,dL_N(x)
  \right]
  =
  0.
  \label{eq:gaussian-log-gas-exponential-ui}
\end{equation}
Hence the same convergence of expectations holds for \(f(x)=\ee^x\).
\end{quotedtheorem}

The last assertion is the point that a saddle calculation usually hides.  A
standard form of the Gaussian edge estimate gives, for \(u\) beyond the
limiting right edge,
\[
  \mathbb P_{\lambda,N}(\lambda_{\max}\ge u)
  \le
  A_\lambda\exp[-N c_\lambda u^2].
\]
Then
\[
  \int_R^\infty \ee^x\,dL_N(x)
  \le
  \ee^{\lambda_{\max}}\mathbf 1_{\lambda_{\max}>R}
\]
and integration by parts bounds the expectation by
\[
  \ee^R\mathbb P(\lambda_{\max}>R)
  +
  \int_R^\infty \ee^u\mathbb P(\lambda_{\max}>u)\,du ,
\]
whose \(N\to\infty\) upper limit vanishes and then remains zero as
\(R\to\infty\).  The left tail is harmless because \(\ee^x\le \ee^{-R}\) on
\((-\infty,-R]\).  Thus one first applies weak convergence to a bounded
truncation of \(\ee^x\), and only then removes the cutoff by
\eqref{eq:gaussian-log-gas-exponential-ui}.  This is the finite-\(N\) bridge
between the matrix integral and the planar Wilson-loop expectation.

\begin{theorem}[Planar circular Wilson loop from the Gaussian model]
\label{thm:planar-circular-wilson-bessel}
Let \(L_N=N^{-1}\sum_i\delta_{x_i}\) for the Gaussian model
\eqref{eq:circular-wilson-eigenvalue-measure}, with
\(\lambda_{\rm circ}>0\) fixed.  Then \(L_N\) converges in probability to
\(\rho_{\lambda_{\rm circ}}(x)\,dx\), and almost surely for an independent
product realization of the finite-\(N\) ensembles, where
\begin{equation}
  \rho_{\lambda_{\rm circ}}(x)
  =
  \frac{2}{\pi\lambda_{\rm circ}}
  \sqrt{\lambda_{\rm circ}-x^2}\,
  \mathbf 1_{[-\sqrt{\lambda_{\rm circ}},\sqrt{\lambda_{\rm circ}}]}(x).
  \label{eq:circular-wilson-semicircle-density}
\end{equation}
For bounded continuous test functions the expectations of
\(\int f\,dL_N\) converge to the corresponding semicircle integral, and the
unbounded Wilson insertion is controlled by
\eqref{eq:gaussian-log-gas-exponential-ui}.  Therefore the planar
fundamental circular Wilson loop is
\begin{equation}
  \lim_{N\to\infty}
  \left\langle
    \frac1N\operatorname{Tr}_{\square}\ee^X
  \right\rangle_{\lambda_{\rm circ},N}
  =
  \int dx\,\rho_{\lambda_{\rm circ}}(x)\ee^x
  =
  \frac{2}{\sqrt{\lambda_{\rm circ}}}
  I_1\!\left(\sqrt{\lambda_{\rm circ}}\right),
  \label{eq:planar-circular-wilson-bessel}
\end{equation}
where \(I_1\) is the modified Bessel function.
\end{theorem}

\begin{proof}
The quoted Gaussian log-gas theorem gives tightness, concentration at the
unique minimizer, and the expectation statement, including the unbounded
exponential insertion.  It remains to identify the minimizer and evaluate its
exponential moment.  For an empirical density \(\rho\) with
\(\int\rho(x)dx=1\), the rate functional obtained from
\eqref{eq:circular-wilson-eigenvalue-measure} is, up to an additive constant,
\begin{equation}
  \mathcal F[\rho]
  =
  \frac{2}{\lambda_{\rm circ}}\int x^2\rho(x)dx
  -
  \int\!\!\int
  \rho(x)\rho(y)\log|x-y|\,dx\,dy .
  \label{eq:circular-wilson-density-functional}
\end{equation}
Varying \(\rho\) under the normalization constraint gives, on the support of
the minimizing density,
\begin{equation}
  \frac{2}{\lambda_{\rm circ}}x^2
  -
  2\int\rho(y)\log|x-y|\,dy
  =
  \text{constant}.
  \label{eq:circular-wilson-euler-lagrange}
\end{equation}
Differentiating inside the support gives the singular-integral equation
\begin{equation}
  \frac{4x}{\lambda_{\rm circ}}
  =
  2\,\operatorname{p.v.}
  \int\frac{\rho(y)}{x-y}\,dy .
  \label{eq:circular-wilson-saddle-equation}
\end{equation}
The Cauchy transform of
\eqref{eq:circular-wilson-semicircle-density} is
\begin{equation}
  G(z)
  =
  \int
  \frac{\rho_{\lambda_{\rm circ}}(y)}{z-y}\,dy
  =
  \frac{2}{\lambda_{\rm circ}}
  \left(z-\sqrt{z^2-\lambda_{\rm circ}}\right),
  \label{eq:circular-wilson-cauchy-transform}
\end{equation}
where the branch is chosen so that \(G(z)\sim z^{-1}\) at infinity.  The
average of the boundary values on the cut is
\[
  \frac12\bigl(G(x+\ii0)+G(x-\ii0)\bigr)
  =
  \frac{2x}{\lambda_{\rm circ}},
\]
which is precisely the principal value integral in
\eqref{eq:circular-wilson-saddle-equation}.  The normalization of
\eqref{eq:circular-wilson-semicircle-density} follows from
\(\int_{-a}^{a}\sqrt{a^2-x^2}\,dx=\pi a^2/2\), with
\(a^2=\lambda_{\rm circ}\).
Because the quadratic potential is strictly convex, the logarithmic-energy
minimizer is unique.  The displayed density satisfies the Euler--Lagrange
equation on its support.  Outside the support, differentiating the difference
between the left side of
\eqref{eq:circular-wilson-euler-lagrange} and its constant value gives
\[
  \frac{4}{\lambda_{\rm circ}}
  \sqrt{x^2-\lambda_{\rm circ}}
  \quad (x>\sqrt{\lambda_{\rm circ}})
\]
and the corresponding negative derivative on the left component
\(x<-\sqrt{\lambda_{\rm circ}}\).  Hence the effective potential is larger
off the support, so the candidate density is the unique equilibrium density
\(\mu_{\lambda_{\rm circ}}\) in
Quoted Theorem~\ref{qt:gaussian-log-gas-exponential-insertion}.

By the exponential-insertion part of
Quoted Theorem~\ref{qt:gaussian-log-gas-exponential-insertion},
\[
  \lim_{N\to\infty}
  \mathbb E_{\lambda_{\rm circ},N}\!\left[\int\ee^x\,dL_N(x)\right]
  =
  \int\ee^x\,\rho_{\lambda_{\rm circ}}(x)\,dx .
\]
Put
\(z=\sqrt{\lambda_{\rm circ}}\) and \(x=zt\).  Then
\begin{equation}
  \int dx\,\rho_{\lambda_{\rm circ}}(x)\ee^x
  =
  \frac{2}{\pi}
  \int_{-1}^{1}\ee^{zt}\sqrt{1-t^2}\,dt .
  \label{eq:circular-wilson-bessel-integral-step}
\end{equation}
The integral representation
\begin{equation}
  I_1(z)
  =
  \frac{z}{\pi}
  \int_{-1}^{1}\ee^{zt}\sqrt{1-t^2}\,dt
  \label{eq:bessel-i1-semicircle-representation}
\end{equation}
gives \eqref{eq:planar-circular-wilson-bessel}.  Equivalently, expanding
\(\ee^x\) and using the even moments
\begin{equation}
  \int dx\,\rho_{\lambda_{\rm circ}}(x)x^{2n}
  =
  \frac{1}{n+1}\binom{2n}{n}
  \left(\frac{\lambda_{\rm circ}}{4}\right)^n
  \label{eq:circular-wilson-semicircle-moments}
\end{equation}
gives the same series
\begin{equation}
  \sum_{n=0}^{\infty}
  \frac1{n!(n+1)!}
  \left(\frac{\lambda_{\rm circ}}{4}\right)^n
  =
  \frac{2}{\sqrt{\lambda_{\rm circ}}}
  I_1\!\left(\sqrt{\lambda_{\rm circ}}\right).
  \label{eq:circular-wilson-bessel-series}
\end{equation}

The exact Laguerre formula
\eqref{eq:finite-n-circular-wilson-laguerre} gives the same expectation limit
without referring to the equilibrium measure.  Indeed,
\[
  \frac1N L_{N-1}^{(1)}
  \!\left(-\frac{\lambda_{\rm circ}}{4N}\right)
  =
  \sum_{k=0}^{N-1}
  \frac{\binom{N}{k+1}}{N^{k+1}}
  \frac1{k!}
  \left(\frac{\lambda_{\rm circ}}4\right)^k ,
\]
the coefficient is bounded by
\(\bigl(k!(k+1)!\bigr)^{-1}(\lambda_{\rm circ}/4)^k\), and for each fixed
\(k\) it converges to that bound as \(N\to\infty\).  Dominated convergence of
this positive series, together with
\(\exp(\lambda_{\rm circ}/8N)\to1\), reproduces
\eqref{eq:circular-wilson-bessel-series}.  This direct finite-\(N\) check is
consistent with, but logically weaker than, the log-gas statement because it
proves only this protected exponential moment rather than convergence of the
matrix-model empirical measure.
\end{proof}

\begin{remark}[Bremsstrahlung derivative algebra]
\label{rem:bremsstrahlung-derivative-algebra}
Assume the protected circular-loop Ward identity
\begin{equation}
  B(\lambda_{\rm circ})
  =
  \frac{1}{2\pi^2}
  \lambda_{\rm circ}\frac{d}{d\lambda_{\rm circ}}
  \log
  \left\langle
    \frac1N\operatorname{Tr}_{\square}\ee^X
  \right\rangle_{\rm planar}.
  \label{eq:bremsstrahlung-circular-loop-ward}
\end{equation}
Then Theorem~\ref{thm:planar-circular-wilson-bessel} gives
\begin{equation}
  B(\lambda_{\rm circ})
  =
  \frac{\sqrt{\lambda_{\rm circ}}}{4\pi^2}
  \frac{I_2(\sqrt{\lambda_{\rm circ}})}
       {I_1(\sqrt{\lambda_{\rm circ}})} .
  \label{eq:bremsstrahlung-bessel-from-loop}
\end{equation}

Let \(z=\sqrt{\lambda_{\rm circ}}\) and
\[
  W_0(z)=\frac{2I_1(z)}{z}.
\]
Since \(\lambda_{\rm circ}\partial_{\lambda_{\rm circ}}
=(z/2)\partial_z\),
\[
  \lambda_{\rm circ}\frac{d}{d\lambda_{\rm circ}}\log W_0
  =
  \frac z2
  \left(
    \frac{I_1'(z)}{I_1(z)}-\frac1z
  \right).
\]
The modified-Bessel recurrence
\begin{equation}
  I_1'(z)=I_2(z)+\frac1z I_1(z)
  \label{eq:bessel-i1-derivative-recurrence}
\end{equation}
therefore gives
\begin{equation}
  \lambda_{\rm circ}\frac{d}{d\lambda_{\rm circ}}\log W_0
  =
  \frac z2\frac{I_2(z)}{I_1(z)} .
  \label{eq:circular-loop-log-derivative-bessel}
\end{equation}
Substitution into \eqref{eq:bremsstrahlung-circular-loop-ward} proves
\eqref{eq:bremsstrahlung-bessel-from-loop}.
\end{remark}

\section{\texorpdfstring
  {\(S^3\) Localization In Three-Dimensional \(\mathcal N=2\) Theories}
  {S3 Localization In Three-Dimensional N=2 Theories}}

Let \(G\), \(T\), \(\mathfrak t\), \(W_G\), and \(\Delta\) be as above.  On
the round \(S^3_r\), the Coulomb-branch localization variable is the constant
real scalar \(\sigma\in\mathfrak t\) in the vector multiplet.  For a general
compact connected gauge group, however, the Chern--Simons input is not a scalar
attached to the Lie algebra.  It is part of the global gauge-theory datum.

\begin{definition}[Round-\(S^3\) Chern--Simons--matter localization datum]
\label{def:s3-cs-matter-localization-datum}
A round-\(S^3_r\) \(\mathcal N=2\) Chern--Simons--matter localization datum is
a package
\[
  \mathfrak D_{S^3}
  =
  \bigl(
    G,\Lambda_{\rm cochar},s,\varphi,\mathcal K_{\rm bare},
    \mathcal R_{\rm PV},\mathcal K_{\rm eff},
    \zeta,\{R_\Phi,m_\Phi,\Delta_\Phi\},C,\mathcal A_\infty
  \bigr).
\]
Here \(G\) is the compact global form of the gauge group,
\(\Lambda_{\rm cochar}=\operatorname{Hom}(U(1),T)\) is its cocharacter
lattice, \(s\) is the spin or non-spin Chern--Simons choice, and \(\varphi\)
is the framing convention.  The bare Chern--Simons coordinate
\(\mathcal K_{\rm bare}\) is an allowed integral symmetric bilinear form on the
gauge and background cocharacter lattice, with gauge-gauge, gauge-flavor,
flavor-flavor, gauge-\(R\), flavor-\(R\), \(R\)-\(R\), and gravitational/framing
contact components.  On a non-spin theory the self-pairings are restricted by
the corresponding bosonic level lattice; on a spin theory odd diagonal
components are allowed only after the spin structure has been made part of the
theory.

The regulator \(\mathcal R_{\rm PV}\) fixes the parity-anomaly convention.  If
\(\psi\) ranges over the dynamical fermions, with Pauli--Villars signs
\(\epsilon_\psi=\pm1\), then the effective contact form that enters the
localized answer is
\begin{equation}
  \mathcal K_{\rm eff}
  =
  \mathcal K_{\rm bare}
  +
  \frac12
  \sum_{\psi}\epsilon_\psi\,\mathcal K_\psi,
  \label{eq:s3-parity-anomaly-effective-level}
\end{equation}
where, on cocharacters \(\lambda,\lambda'\) and background cocharacters
\(\nu,\nu'\),
\[
  \mathcal K_\psi((\lambda,\nu),(\lambda',\nu'))
  =
  \sum_{\rho\in{\rm wt}(R_\psi)}
  \bigl(\rho(\lambda)+q_\psi(\nu)\bigr)
  \bigl(\rho(\lambda')+q_\psi(\nu')\bigr).
\]
The parity-anomaly condition is that
\(\mathcal K_{\rm eff}\) belongs to the allowed spin or non-spin level lattice.
Changing \(\mathcal R_{\rm PV}\) shifts \(\mathcal K_{\rm bare}\) and the
background contact terms by integral counterterms; it does not leave a
regulator-independent scalar level.  The FI parameter is restricted to
\begin{equation}
  \zeta\in\operatorname{Hom}(G,U(1))\otimes_\mathbb Z\mathbb R
  \simeq
  (\mathfrak t/[\mathfrak g,\mathfrak g])^\ast,
  \label{eq:s3-fi-abelian-center}
\end{equation}
so it vanishes on semisimple coroot directions.  Finally \(C\) is the Cartan
integration cycle and \(\mathcal A_\infty\) is the large-\(|\sigma|\) and pole
prescription used to make the oscillatory integral a regulated localization
functional.
\end{definition}

Let \(m_B\) collect all background real masses and \(R\)-background coordinates
that enter the contact terms, and write the gauge-background block of
\(\mathcal K_{\rm eff}\) as \(\mathcal K_{\rm eff}^{gB}\).  The
Kapustin--Willett--Yaakov Cartan integral attached to
\(\mathfrak D_{S^3}\) is
\begin{equation}
\begin{aligned}
  Z_{S^3}(\mathfrak D_{S^3})
  &=
  \frac1{|W_G|}
  \int_C d\sigma\,
  \exp\!\bigl[\mathcal I_{\rm CS}(\sigma;m_B)\bigr]\\
  &\quad\times
  \prod_{\alpha\in\Delta_+}
  4\sinh^2\!\bigl(\pi r\,\alpha(\sigma)\bigr)
  \prod_{\Phi}Z_{\Phi}^{S^3}(\sigma;m_B),
\end{aligned}
  \label{eq:kwy-s3-localization}
\end{equation}
where the contact phase is
\begin{equation}
  \begin{aligned}
  \mathcal I_{\rm CS}(\sigma;m_B)
  &=
  \ii\pi\mathcal K_{\rm eff}^{gg}(r\sigma,r\sigma)
  +2\pi\ii\mathcal K_{\rm eff}^{gB}(r\sigma,m_B)\\
  &\quad
  +\ii\pi\mathcal K_{\rm eff}^{BB}(m_B,m_B)
  +2\pi\ii r\langle\zeta,\sigma\rangle .
  \end{aligned}
  \label{eq:s3-localization-contact-phase}
\end{equation}
The chiral factor is
\[
  Z_{\Phi}^{S^3}(\sigma;m_B)
  =
  \prod_{\rho\in{\rm wt}(R_\Phi)}
  \exp\!\left[
    \ell\!\left(
      1-\Delta_\Phi
      +\ii r\,\rho(\sigma)
      +\ii\,q_\Phi(m_B)
    \right)
  \right].
\]
Thus the displayed integral is a function of the full contact-level datum.  The
single symbol \(k\) is recovered only in specializations where the global form,
spin choice, trace normalization, regulator, and background contact scheme have
collapsed \(\mathcal K_{\rm eff}^{gg}\) to one integral multiple of a basic
quadratic form.
The function \(\ell\) is fixed, up to an additive constant that changes the
overall normalization of \(Z_{S^3}\), by
\begin{equation}
  \ell'(z)=-\pi z\cot(\pi z),
  \qquad
  \ell\!\left(\frac12+\ii x\right)
  +
  \ell\!\left(\frac12-\ii x\right)
  =
  -\log\!\bigl(2\cosh(\pi x)\bigr).
  \label{eq:kwy-ell-identity}
\end{equation}
Equivalently, the round-sphere determinant is the \(b=1\) specialization of
the double-sine convention
\begin{equation}
  s_b(x)
  =
  \prod_{m,n\ge0}
  \frac{mb+nb^{-1}+\frac{Q_b}{2}-\ii x}
       {mb+nb^{-1}+\frac{Q_b}{2}+\ii x},
  \qquad Q_b=b+b^{-1},
  \label{eq:double-sine-definition}
\end{equation}
defined by zeta regularization or by the corresponding meromorphic
continuation.  In the convention of \eqref{eq:kwy-s3-localization},
\begin{equation}
  \exp\!\left[\ell(1-\Delta+\ii x)\right]
  =
  s_1\!\left(\ii(1-\Delta)-x\right),
  \qquad x\in\mathbb R,
  \label{eq:ell-double-sine-round}
\end{equation}
after fixing the same \(x\)-independent normalization on both sides.  The
reflection identity used for a conjugate \(R\)-charge \(1/2\) pair is
\begin{equation}
  s_1\!\left(\frac{\ii}{2}-x\right)
  s_1\!\left(\frac{\ii}{2}+x\right)
  =
  \frac1{2\cosh(\pi x)} .
  \label{eq:double-sine-hyper-identity}
\end{equation}

The integration contour is the real Cartan in the displayed unitary
normalization only after \(\mathcal K_{\rm eff}\), the FI coordinate, and the
matter determinants have been fixed.  Along a ray \(\sigma=t v\), the large
\(t\) test uses the full exponent
\[
  \ii\pi t^2\mathcal K_{\rm eff}^{gg}(rv,rv)
  +2\pi\ii t\,
  \bigl(\mathcal K_{\rm eff}^{gB}(rv,m_B)
        +r\langle\zeta,v\rangle\bigr)
  +
  \sum_{\Phi,\rho}
  \log Z_{\ell,\Phi}\!\left(t\,\rho(v)\right)
\]
together with the vector-multiplet hyperbolic-sine growth.  If this asymptotic
behavior is not damping in every Weyl chamber, then a real mass, FI chamber,
analytic continuation, or residue prescription must be declared as part of the
theory being evaluated.  In abelianized descriptions the possible poles lie on
complexified hyperplanes
\[
  \rho(\sigma)+m_\Phi
  +\ii\bigl(n+\Delta_\Phi\bigr)/r=0,
  \qquad n\in\mathbb Z_{\ge0},
\]
up to the convention used for \(\ell\) or \(s_b\).  A Jeffrey--Kirwan-type
residue prescription is therefore not a decorative evaluation trick: it is a
choice of chamber and integration cycle for the regulated localization structure.

\begin{example}[Simply connected simple specialization]
\label{ex:s3-simple-simply-connected-cs-level}
If \(G\) is simple and simply connected, and no background contact terms are
turned on, the gauge-gauge block can be written
\[
  \mathcal K_{\rm eff}^{gg}(\sigma,\sigma)
  =
  k_{\rm eff}(\sigma,\sigma)_{\rm basic},
  \qquad k_{\rm eff}\in\mathbb Z,
\]
in the basic integral normalization.  There is no FI parameter for a semisimple
group.  This is the common setting in which the formula looks as if it had one
integer \(k_{\rm eff}\).  For a quotient global form or a non-spin theory the
allowed values are instead the corresponding sublattice of integral
Chern--Simons forms.
\end{example}

\begin{example}[Standard \(U(N)\) factor]
\label{ex:s3-un-standard-level-data}
For the standard \(U(N)\) factor used in many localization formulae, choose
\(\Lambda_{\rm cochar}\simeq\mathbb Z^N\) with Weyl group \(S_N\).  The diagonal
level \(U(N)_k\) has
\[
  \mathcal K_{\rm eff}^{gg}(\sigma,\sigma)
  =
  k\sum_{i=1}^N\sigma_i^2,
  \qquad k\in\mathbb Z
\]
on spin \(S^3\).  A non-spin oriented theory imposes the corresponding even
self-pairing condition.  The FI coordinate is necessarily central:
\[
  \langle\zeta,\sigma\rangle
  =
  \zeta\sum_{i=1}^N\sigma_i .
\]
With chiral fermions present, the integer \(k\) in this display is the
effective level after the parity-anomaly shift
\[
  \mathcal K_{\rm eff}^{gg}(\lambda,\lambda')
  =
  k_{\rm bare}\sum_i\lambda_i\lambda_i'
  +
  \frac12
  \sum_\psi\epsilon_\psi
  \sum_{\rho\in{\rm wt}(R_\psi)}
  \rho(\lambda)\rho(\lambda')
\]
has landed in the allowed \(U(N)\) level lattice.
\end{example}

\begin{example}[Product group and mixed determinant levels]
\label{ex:s3-product-group-mixed-levels}
For \(U(N)\times U(M)\), with Cartan variables
\((\sigma_i,\widehat\sigma_j)\), the standard opposite-level gauge block is
\[
  \mathcal K_{\rm eff}^{gg}
  =
  k\sum_{i=1}^N\sigma_i^2
  -
  k\sum_{j=1}^M\widehat\sigma_j^2 .
\]
The most general Abelian determinant-sector contact datum may also include
mixed integral terms such as
\[
  2h
  \left(\sum_i\sigma_i\right)
  \left(\sum_j\widehat\sigma_j\right),
  \qquad h\in\mathbb Z,
\]
subject again to the spin or non-spin level lattice.  A bifundamental chiral
fermion has weights \(e_i-\widehat e_j\), so its parity-anomaly contribution is
added to the two diagonal blocks and to the mixed block before the matrix model
is declared.  The ABJM specialization of
Chapter~\ref{ch:susy-three-dimensional-chern-simons-matter} sets the standard
unquotiented global form, opposite levels, bifundamental matter,
spin/framing convention, line lattice, and background contact scheme first; only
then does the familiar two-matrix \(S^3\) integral have a QFT meaning.
\end{example}

\begin{example}[Abelian Chern--Simons Gaussian]
\label{ex:s3-u1-cs-fresnel}
For spin pure \(U(1)_k\) Chern--Simons theory on \(S^3\), \(k\in\mathbb Z\);
for the non-spin oriented version in this normalization, the diagonal
self-level is even.  With FI parameter \(\zeta\), which is allowed because
\(U(1)\) is Abelian, the localized integral is the oscillatory Fresnel integral
\[
  Z_{U(1)_k}(\zeta)
  =
  \int_{\mathbb R} d\sigma\,
  \exp\!\left(\ii\pi k\sigma^2+2\pi\ii\zeta\sigma\right).
\]
With the contour obtained by rotating the positive quadratic direction by the
usual \(\ii0\) prescription, completing the square gives
\begin{equation}
  Z_{U(1)_k}(\zeta)
  =
  \frac{\ee^{\ii\pi\,{\rm sgn}(k)/4}}{\sqrt{|k|}}\,
  \exp\!\left(-\frac{\ii\pi\zeta^2}{k}\right).
  \label{eq:s3-u1-cs-fresnel}
\end{equation}
The phase depends on the framing convention of the Chern--Simons theory; the
absolute value and FI dependence follow from the declared contour.
\end{example}

\begin{example}[A conjugate pair of chirals]
\label{ex:s3-chiral-pair-integral}
For two chiral multiplets of charges \(+1\) and \(-1\), both with
\(R\)-charge \(1/2\) and zero real mass, the product of one-loop determinants
on the \(U(1)\) Coulomb parameter is
\[
  \exp\!\left[
    \ell\!\left(\frac12+\ii\sigma\right)
    +
    \ell\!\left(\frac12-\ii\sigma\right)
  \right]
  =
  \frac1{2\cosh(\pi\sigma)}.
\]
The elementary integral
\begin{equation}
  \int_{-\infty}^{\infty}
  \frac{d\sigma}{2\cosh(\pi\sigma)}
  =
  \frac12
  \label{eq:s3-chiral-pair-integral}
\end{equation}
is a finite check on the determinant normalization in
\eqref{eq:kwy-s3-localization}.  It is not by itself a definition of a gauge
theory; the gauge quotient, possible Chern--Simons contact terms, and
background-field couplings remain part of the QFT definition.
\end{example}

\section{Theorem Boundary And Singular Loci}

Equations~\eqref{eq:pestun-s4-localization} and
\eqref{eq:kwy-s3-localization} have different mathematical status from
Remark~\ref{rem:susy-localization-deformation-identity}.  That
remark records the finite-dimensional deformation identity under stated
hypotheses.  The compact-space QFT formulae additionally require:
\[
\begin{array}{ll}
S^4:&
\text{a regulator preserving the chosen supercharge, determinant
regularization,}\\
&\text{and an instanton compactification or contour for singular point
instantons;}\\[3pt]
S^3:&
\text{a global Chern--Simons/contact level, parity-anomaly regulator,}\\
&\text{framing convention, real contour or analytic continuation,}\\
&\text{and a treatment of noncompact Coulomb directions and singular
hyperplanes.}
\end{array}
\]
When these data are supplied, localization turns a protected observable into
an integral over fixed-locus variables.  It does not erase the distinction
between regularization and renormalization, nor does it make contact terms
canonical.  Background Chern--Simons terms in three dimensions and local
curvature counterterms in four dimensions change phases or local
source-dependent factors in the localized answer; they must be tracked as
coordinates of the QFT, not discarded as normalization details.
