\chapter{Mirror TBA and the Y-System}
\label{ch:planar-n4-mirror-tba-y-system}

\ConventionsLorentzian

The asymptotic Bethe ansatz determines long-chain states before wrapping
effects.  Finite single traces require a finite-size formalism.  The planar
\(\mathcal N=4\) construction uses a mirror model obtained by exchanging
worldsheet space and Euclidean time in the integrable spin-chain scattering
problem.  This chapter states the logical structure without pretending that
the mirror thermodynamic Bethe ansatz is a theorem of four-dimensional
constructive QFT.

\begin{warning}[Status of mirror magnons]
\label{warn:planar-n4-mirror-magnon-status}
A mirror magnon is not a new physical particle of four-dimensional
\(\mathcal N=4\) SYM.  It is the excitation obtained after analytically
continuing the two-dimensional worldsheet/spin-chain kinematics so that
Euclidean time and space are exchanged.  In the relativistic string
worldsheet before light-cone gauge fixing this exchange is the natural
mirror-channel operation.  After gauge fixing, the magnon dispersion is
nonrelativistic and the same operation becomes a path on the Zhukovsky
surface together with a specified dressing-phase branch.  This is why the
mirror relation is not explained by the nonrelativistic spin chain alone.

The reason to use mirror magnons in the gauge-theory spectral problem is
therefore indirect.  Once the long-trace magnon S-matrix is accepted as the
asymptotic integrable description, finite trace length is computed by the
standard mirror-channel finite-size construction: virtual mirror particles
wrap around the trace, and their thermodynamics gives the TBA.  This is the
worldsheet and integrability explanation.  There is no known purely
four-dimensional derivation in which the mirror magnon appears directly as an
ordinary local QFT asymptotic state.
\end{warning}

\section{Mirror Transformation}

For a physical magnon with energy \(E(p)\) and momentum \(p\), the mirror
variables are defined by the stringbook analytic continuation
\[
  E=\ii\widetilde p,\qquad
  p=\ii\widetilde E ,
  \qquad\text{equivalently}\qquad
  \widetilde E=-\ii p,\quad
  \widetilde p=-\ii E .
\]
Since
\[
  E(p)=\sqrt{1+16g^2\sin^2\frac p2},
\]
the mirror dispersion is not the same function of \(\widetilde p\) as the
physical dispersion.  This is the first major difference from relativistic
two-dimensional QFT, where the mirror transformation is a rapidity shift.

\begin{proposition}[Mirror dispersion from double Wick rotation]
\label{prop:planar-n4-mirror-dispersion}
For a \(Q\)-magnon bound state with physical dispersion
\[
  E_Q(p)=\sqrt{Q^2+16g^2\sin^2\frac p2},
\]
the mirror dispersion obtained from
\(p=\ii\widetilde E\), \(E_Q=\ii\widetilde p\) is
\begin{equation}
  \widetilde E_Q(\widetilde p)
  =
  2\,\operatorname{arsinh}
  \left(
    \frac{\sqrt{Q^2+\widetilde p^{\,2}}}{4g}
  \right).
  \label{eq:planar-n4-mirror-dispersion}
\end{equation}
The positive branch is chosen for real mirror momentum.
\end{proposition}

\begin{proof}
Set \(p=\ii\widetilde E\) and \(E_Q=\ii\widetilde p\) in the physical
dispersion relation.  Since
\(\sin(\ii\widetilde E/2)=\ii\sinh(\widetilde E/2)\), one obtains
\[
  -\widetilde p^{\,2}
  =
  Q^2-16g^2\sinh^2\frac{\widetilde E}{2}.
\]
Thus
\[
  \sinh^2\frac{\widetilde E}{2}
  =
  \frac{Q^2+\widetilde p^{\,2}}{16g^2}.
\]
The mirror energy is positive on the real mirror line, which selects
\eqref{eq:planar-n4-mirror-dispersion}.
\end{proof}

Equivalently one continues the Zhukovsky variables through a specified
sequence of cuts.  On the physical sheet a real magnon has
\((x^+)^*=x^-\) and \(|x^\pm|>1\).  A mirror particle of real mirror rapidity
has one Zhukovsky variable inside and one outside the unit circle.  The
choice of path is part of the definition of the mirror S-matrix: crossing a
different cut changes the scalar dressing factor by the crossing equation.
For this reason the mirror TBA below always includes the contour and sheet
data together with the functional formulae.

It is useful to keep the following sheet ledger explicit.  Let
\(x^{[\pm Q]}(u)=x(u\pm\ii Q/2)\), with \(x(u)\) the physical Zhukovsky
branch of Chapter~\ref{ch:planar-n4-asymptotic-bethe-ansatz}.  Physical
bound-state kinematics keeps both \(x^{[+Q]}\) and \(x^{[-Q]}\) on the
outside sheet for real momentum.  Mirror kinematics continues one of the two
variables through the short cut, so the pair has opposite inside/outside
status.  The transfer matrix and the dressing factor must be continued along
the same path.  A formula copied from the physical sheet but evaluated with
the wrong inside/outside assignment gives the wrong wrapping sign.

Bound states of \(Q\) physical magnons are described by pairs
\[
  x_Q^+,\qquad x_Q^-,
\]
obeying the bound-state shortening relation
\[
  x_Q^++\frac1{x_Q^+}-x_Q^- -\frac1{x_Q^-}
  =
  \frac{\ii Q}{g}.
\]
In the mirror region the branch is chosen so that the mirror energy
\(\widetilde E_Q(u)\) is positive for real mirror rapidity \(u\).

The preceding paragraph is often the place where sign mistakes enter.  The
following elementary parametrization fixes the sheet convention used in the
rest of the chapter and in the stringbook mirror Bethe--Yang equations.

\begin{proposition}[Mirror Zhukovsky sheet parametrization]
\label{prop:planar-n4-mirror-zhukovsky-sheet-parametrization}
Fix \(g>0\), \(Q\in\mathbb Z_{>0}\), and real mirror momentum
\(\widetilde p\).  Put
\[
  R_Q=\sqrt{Q^2+\widetilde p^{\,2}},
  \qquad
  \widetilde E_Q
  =
  2\,\operatorname{arsinh}\frac{R_Q}{4g},
  \qquad
  r_Q=\ee^{-\widetilde E_Q/2},
  \qquad
  \xi_Q=\frac{\widetilde p-\ii Q}{R_Q}.
\]
Then
\begin{equation}
  x_Q^+=r_Q\xi_Q,\qquad
  x_Q^-=r_Q^{-1}\xi_Q
  \label{eq:planar-n4-mirror-zhukovsky-parametrization}
\end{equation}
satisfy
\begin{align}
  x_Q^++\frac1{x_Q^+}-x_Q^- -\frac1{x_Q^-}
  &=
  \frac{\ii Q}{g},
  \label{eq:planar-n4-mirror-shortening-check}\\
  \log\frac{x_Q^-}{x_Q^+}
  &=
  \widetilde E_Q,
  \label{eq:planar-n4-mirror-energy-log-check}\\
  \ii\widetilde p
  &=
  -Q-2\ii g(x_Q^+-x_Q^-).
  \label{eq:planar-n4-mirror-momentum-check}
\end{align}
Moreover \(|x_Q^+|<1<|x_Q^-|\).  At weak coupling with \(Q,\widetilde p\)
fixed,
\begin{equation}
  \ee^{-L\widetilde E_Q}
  =
  \left(\frac{4g^2}{Q^2+\widetilde p^{\,2}}\right)^L
  \bigl(1+O(g^2)\bigr),
  \label{eq:planar-n4-mirror-weak-boltzmann}
\end{equation}
so one mirror wrapping around a length-\(L\) trace first appears at order
\(g^{2L}\) before transfer-matrix and dressing factors are included.
\end{proposition}

\begin{proof}
The number \(\xi_Q\) has unit modulus and \(0<r_Q<1\).  Hence
\(|x_Q^+|<1<|x_Q^-|\), and
\(\log(x_Q^-/x_Q^+)=-2\log r_Q=\widetilde E_Q\) on the real mirror branch.
Since
\[
  \sinh\frac{\widetilde E_Q}{2}=\frac{R_Q}{4g},
  \qquad
  \xi_Q^{-1}-\xi_Q=\frac{2\ii Q}{R_Q},
\]
we get
\[
  x_Q^++\frac1{x_Q^+}-x_Q^- -\frac1{x_Q^-}
  =
  2\sinh\frac{\widetilde E_Q}{2}\,(\xi_Q^{-1}-\xi_Q)
  =
  \frac{\ii Q}{g}.
  \]
Similarly,
\[
  -Q-2\ii g(x_Q^+-x_Q^-)
  =
  -Q+4\ii g\sinh\frac{\widetilde E_Q}{2}\,\xi_Q
  =
  -Q+\ii R_Q\xi_Q
  =
  \ii\widetilde p.
\]
Finally,
\(\ee^{-\widetilde E_Q}=(\sqrt{1+R_Q^2/(16g^2)}-R_Q/(4g))^2\), whose
small-\(g\) expansion gives \eqref{eq:planar-n4-mirror-weak-boltzmann}.
\end{proof}

The same mirror sheet parametrization fixes the large-rapidity fermionic-node
ratio used later to identify the \(Y_{1,1}Y_{2,2}\) energy exponent in the QSC
bridge; see Proposition~\ref{prop:qsc-fermionic-node-large-u-ratio}.  This is a
good example of the convention burden in this subject: the sign comes from the
inside/outside assignment just proved here, but the ratio itself belongs to the
later TBA-to-QSC asymptotic bridge rather than to the definition of the mirror
kernel.

\begin{definition}[Mirror-kernel datum]
\label{def:planar-n4-mirror-kernel-datum}
A mirror kernel in this chapter means the triple
\[
  (S^{\rm mirror}_{BA}(v,u),\mathcal C_B,\mathcal C_A),
\]
where \(S^{\rm mirror}_{BA}\) is the analytically continued two-body
scattering factor and \(\mathcal C_B,\mathcal C_A\) are the source and target
rapidity contours on the chosen sheets.  The derivative kernel is then
\[
  K_{BA}(v,u)
  =
  \frac{1}{2\pi\ii}\partial_v\log S^{\rm mirror}_{BA}(v,u).
  \]
This convention packages the branch choice with the formula and prevents
the common mistake of treating mirror TBA kernels as functions on a single
unbranched \(u\)-plane.
\end{definition}

\section{Mirror Bethe-String Species}

The mirror TBA is nested because a physical magnon carries two
\(\mathfrak{su}(2|2)\) wings of polarization data.  The canonical node
families are as follows.
\[
\begin{array}{c|c|c}
\text{symbol} & \text{range} & \text{role}\\
\hline
\bullet_Q & Q=1,2,\ldots & \text{mirror bound states carrying energy}\\
y_\pm & \text{two sheets} & \text{fermionic auxiliary roots}\\
(v|M) & M=1,2,\ldots & \text{auxiliary strings in each wing}\\
(w|M) & M=1,2,\ldots & \text{isotopic auxiliary strings in each wing}
\end{array}
\]
Each wing has its own copy of the \(v\)- and \(w\)-string nodes.  The
\(\bullet_Q\) nodes are the only nodes that enter the energy formula with the
measure \(\dd\widetilde p_Q/\dd u\).  The auxiliary nodes enter through
kernels and chemical potentials and enforce the nested Bethe constraints.

The string hypothesis used here is a thermodynamic statement on the mirror
Bethe roots: in a large mirror volume, complex roots organize into regular
arrays whose centers become real variables \(u\).  In a relativistic model
this statement is already subtle; in planar \(\mathcal N=4\) it is further
entangled with the Zhukovsky cuts.  Consequently, the node table is part of
the integrability construction, not a theorem about the original
four-dimensional path integral.

The node table above abbreviates arrays selected by the Bethe-Yang
equations.  We now derive the auxiliary arrays from the large-volume
pole-cancellation mechanism in one
\(\mathfrak{su}(2|2)_{\rm c}\) wing.  Put
\[
  \eta=\frac{\ii}{2g},
  \qquad
  v(y)=y+\frac1y .
\]
Let \(x_j^\pm\), \(j=1,\ldots,K^{\rm I}\), be mirror level-I rapidity data
on the real mirror line, so that
\[
  x_j^+=r_j\xi_j,\qquad x_j^-=r_j^{-1}\xi_j,
  \qquad 0<r_j<1,\quad \lvert\xi_j\rvert=1 .
\]
The one-wing auxiliary Bethe-Yang subsystem in the stringbook mirror
orientation has the form
\begin{equation}
\begin{aligned}
  -1
  &=
  F(y_a)
  \prod_{b=1}^{K^{\rm III}}
  \frac{w_b-v(y_a)-\eta}{w_b-v(y_a)+\eta},
  \qquad
  F(y)=
  \prod_{j=1}^{K^{\rm I}}
  \frac{y-x_j^+}{y-x_j^-}
  \sqrt{\frac{x_j^-}{x_j^+}},
  \\
  1
  &=
  \prod_{a=1}^{K^{\rm II}}
  \frac{w_b-v(y_a)+\eta}{w_b-v(y_a)-\eta}
  \prod_{\substack{c=1\\c\neq b}}^{K^{\rm III}}
  \frac{w_c-w_b+2\eta}{w_c-w_b-2\eta}.
\end{aligned}
  \label{eq:planar-n4-mirror-auxiliary-bethe-yang}
\end{equation}
The sign in the first equation is the mirror fermion boundary-condition sign.
It is not the graded sign in the physical nested ABA of
Chapter~\ref{ch:planar-n4-asymptotic-bethe-ansatz}.

\begin{proposition}[Mirror auxiliary support and Bethe strings]
\label{prop:planar-n4-mirror-auxiliary-strings}
Assume the mirror Bethe-Yang equations
\eqref{eq:planar-n4-mirror-auxiliary-bethe-yang}, the mirror sheet data
above, and a thermodynamic limit in which the logarithmic averages of
\(\log\lvert F(y)\rvert\) have the signs determined by the one-particle
factor below.  Then elementary \(y\)-roots have finite density only on
\(\lvert y\rvert=1\), elementary \(w\)-roots have finite density only on the
real \(w\)-axis, and the non-elementary auxiliary complexes have the
following pole-cancellation arrays:
\[
\begin{aligned}
  v(y_{j,{\rm in}})&=v_0+(M+2-2j)\eta,
  &
  \lvert y_{j,{\rm in}}\rvert&<1,
  \\
  v(y_{j,{\rm out}})&=v_0-(M+2-2j)\eta,
  &
  \lvert y_{j,{\rm out}}\rvert&>1,
  \\
  w_j&=v_0+(M+1-2j)\eta,
  &&
  j=1,\ldots,M,
\end{aligned}
\]
with real thermodynamic center \(v_0\).  These are the \((v|M)\), or
\(M|yw\), strings.  The pure \((w|M)\) strings are
\[
  w_j=w_0+(M+1-2j)\eta,
  \qquad j=1,\ldots,M,
\]
with real center \(w_0\).  The assertion is a statement about the mirror
string hypothesis and the displayed large-volume equations; it is not a
theorem of four-dimensional \(\mathcal N=4\) SYM.
\end{proposition}

\begin{proof}
First isolate the source of exponential growth and decay.  For a single
mirror level-I root write \(x^+=r\xi\), \(x^-=r^{-1}\xi\), with
\(0<r<1\) and \(\lvert\xi\rvert=1\).  The magnitude of the one-particle
factor
\[
  s(y;x^\pm)=\frac{y-x^+}{y-x^-}\sqrt{\frac{x^-}{x^+}}
\]
obeys
\[
  \lvert s(y;x^\pm)\rvert^2
  =
  \frac{\lvert y-r\xi\rvert^2}{\lvert ry-\xi\rvert^2},
  \qquad
  \lvert ry-\xi\rvert^2-\lvert y-r\xi\rvert^2
  =
  (1-r^2)(1-\lvert y\rvert^2).
\]
Thus \(\lvert s\rvert<1\) for \(\lvert y\rvert<1\), \(\lvert s\rvert=1\)
for \(\lvert y\rvert=1\), and \(\lvert s\rvert>1\) for
\(\lvert y\rvert>1\).  Multiplying over \(K^{\rm I}\) level-I roots gives
the stated support of elementary \(y\)-roots in the thermodynamic limit,
provided the level-I density has a nonzero logarithmic average.

Now take a \(y\)-root with \(\lvert y_1\rvert<1\).  The factor \(F(y_1)\)
then tends to zero exponentially, so the first equation in
\eqref{eq:planar-n4-mirror-auxiliary-bethe-yang} can remain finite only if
one of the \(w\)-factors has a pole:
\[
  w_1-v(y_1)+\eta=0,
  \qquad\text{that is}\qquad
  w_1=v(y_1)-\eta .
\]
The corresponding \(w_1\)-equation has a zero from the same \(y_1\).  The
local cancellation recursion remains finite by acquiring a pole from a second
\(y\)-root,
\[
  w_1-v(y_2)-\eta=0,
  \qquad
  v(y_2)=v(y_1)-2\eta.
\]
If this second root lies outside the unit circle, the divergence of
\(F(y_2)\) is canceled by the zero of the same \(w_1\)-factor and the string
closes.  If it lies inside, the same argument continues and creates the next
\(w\)-root.  Iteration fixes the displayed imaginary spacings.  The real
center is the thermodynamic string-hypothesis input: the mirror densities are
assigned to real centers, while the roots in the complex plane are the
regular arrays above those centers.

The \(w\)-only strings are obtained by applying the same valuation argument to
the second line of \eqref{eq:planar-n4-mirror-auxiliary-bethe-yang} when the
background \(y\)-roots are on \(\lvert y\rvert=1\).  A finite elementary
density is supported on real \(w\).  A complex \(w\)-root must be accompanied
by neighbouring \(w\)-roots separated by \(2\eta=\ii/g\), which gives the
array \(w_j=w_0+(M+1-2j)\eta\).
\end{proof}

\section{Thermodynamic Bethe Ansatz}

The mirror TBA is a thermodynamic limit of Bethe-Yang quantization, not an
independent postulate.  We first derive the one-species formula because it is
the place where the signs, statistics, and kernel orientation are least
forgiving.  The derivation is also the convention bridge to the stringbook:
the continuum density equation naturally differentiates the scattering phase in
the target rapidity, while the multi-species formula below is written with the
source-derivative kernels of
Definition~\ref{def:planar-n4-mirror-kernel-datum}.

\begin{proposition}[One-species mirror TBA from Bethe-Yang quantization]
\label{prop:one-species-mirror-tba-derivation}
Consider a diagonal integrable mirror model with one particle species,
real mirror rapidity \(u\), mirror momentum \(\widetilde p(u)\), mirror
energy \(\widetilde\epsilon(u)\), and two-body scattering factor
\(S(v,u)\).  Assume:
\begin{enumerate}
\item a logarithm of \(S(v,u)\) has been chosen with no self-scattering jump at
  \(v=u\);
\item the thermodynamic mode number \(n(u)\) is smooth and monotone on the
  rapidity interval under consideration;
\item the finite-volume Bethe levels have Fermi level statistics, so a level is
  either occupied or empty;
\item no additional Bethe-string species or bound-state poles cross the
  integration contour in the ground-state calculation.
\end{enumerate}
Let \(\rho_{\rm p}\), \(\rho_{\rm L}\), and
\(\rho_{\rm h}=\rho_{\rm L}-\rho_{\rm p}\) denote the particle, level, and hole
densities with respect to \(u\).  Define the density kernel
\begin{equation}
  K^{\rm dens}(v,u)
  =
  \frac{1}{2\pi\ii}\partial_u\log S(v,u).
  \label{eq:one-species-density-kernel}
\end{equation}
Then the thermodynamic limit of the Bethe-Yang equation gives
\begin{equation}
  \rho_{\rm L}(u)
  =
  \frac{\widetilde L}{2\pi}
  \frac{\dd \widetilde p(u)}{\dd u}
  -
  \int\dd v\,\rho_{\rm p}(v)K^{\rm dens}(v,u).
  \label{eq:one-species-density-equation}
\end{equation}
At inverse mirror temperature \(\widetilde\beta\) and mirror chemical potential
\(\widetilde\mu\), extremizing
\[
  \mathfrak s
  -
  \widetilde\beta\int\dd u\,\rho_{\rm p}(u)\widetilde\epsilon(u)
  -
  \widetilde\mu\int\dd u\,\rho_{\rm p}(u)
\]
subject to \eqref{eq:one-species-density-equation}, where
\[
  \mathfrak s
  =
  \int\dd u\,
  \left[
    \rho_{\rm L}\log\rho_{\rm L}
    -
    \rho_{\rm p}\log\rho_{\rm p}
    -
    \rho_{\rm h}\log\rho_{\rm h}
  \right],
\]
gives the pseudoenergy equation
\begin{equation}
  \zeta(u)
  =
  \widetilde\beta\,\widetilde\epsilon(u)
  +
  \widetilde\mu
  +
  \int\dd v\,K^{\rm dens}(u,v)\log(1+\ee^{-\zeta(v)}),
  \qquad
  \frac{\rho_{\rm p}(u)}{\rho_{\rm L}(u)}
  =
  \frac{1}{1+\ee^{\zeta(u)}} .
  \label{eq:one-species-tba}
\end{equation}
At the stationary point,
\begin{equation}
  \log Z
  =
  \widetilde L
  \int\frac{\dd u}{2\pi}\,
  \frac{\dd \widetilde p(u)}{\dd u}
  \log(1+\ee^{-\zeta(u)}).
  \label{eq:one-species-log-z}
\end{equation}
After exchanging mirror space and Euclidean time, with
\(\widetilde\beta=L\) and no chemical potential, this gives
\begin{equation}
  E_0(L)
  =
  -
  \int\frac{\dd u}{2\pi}\,
  \frac{\dd \widetilde p(u)}{\dd u}
  \log(1+\ee^{-\zeta(u)}).
  \label{eq:one-species-ground-state-energy}
\end{equation}
\end{proposition}

\begin{proof}
For occupied rapidities \(u_k\), choose the logarithmic Bethe-Yang equation in
the stringbook orientation
\[
  \widetilde p(u_k)\widetilde L
  =
  2\pi n_k
  -
  \ii\sum_{j\neq k}\log S(u_j,u_k).
\]
Replacing the sum by
\(\int\dd v\,\rho_{\rm p}(v)\) and differentiating with respect to the target
rapidity \(u\) gives
\[
  \widetilde L\,\widetilde p'(u)
  =
  2\pi\rho_{\rm L}(u)
  +
  2\pi\int\dd v\,\rho_{\rm p}(v)K^{\rm dens}(v,u),
\]
which is \eqref{eq:one-species-density-equation}.

Now vary the entropy while preserving the density equation.  Since
\(\rho_{\rm h}=\rho_{\rm L}-\rho_{\rm p}\),
\[
  \delta\mathfrak s
  =
  \int\dd u\,
  \left[
    \log\frac{\rho_{\rm L}}{\rho_{\rm h}}\delta\rho_{\rm L}
    -
    \log\frac{\rho_{\rm p}}{\rho_{\rm h}}\delta\rho_{\rm p}
  \right].
\]
The constraint implies
\[
  \delta\rho_{\rm L}(u)
  =
  -
  \int\dd v\,K^{\rm dens}(v,u)\delta\rho_{\rm p}(v).
\]
Collecting the coefficient of an arbitrary
\(\delta\rho_{\rm p}(u)\) gives
\[
  \log\frac{\rho_{\rm h}(u)}{\rho_{\rm p}(u)}
  =
  \widetilde\beta\,\widetilde\epsilon(u)+\widetilde\mu
  +
  \int\dd v\,K^{\rm dens}(u,v)
    \log\frac{\rho_{\rm L}(v)}{\rho_{\rm h}(v)} .
\]
With
\(\zeta=\log(\rho_{\rm h}/\rho_{\rm p})\), one has
\(\rho_{\rm p}/\rho_{\rm L}=1/(1+\ee^\zeta)\) and
\(\rho_{\rm L}/\rho_{\rm h}=1+\ee^{-\zeta}\), proving
\eqref{eq:one-species-tba}.

It remains to evaluate the extremized functional.  Substitute the stationary
equation into the expression for
\(\mathfrak s-\widetilde\beta\widetilde E-\widetilde\mu\widetilde N\).  The
terms proportional to \(\rho_{\rm p}\) cancel against the convolution term in
the entropy variation, and the remaining density equation leaves
\[
  \log Z
  =
  \int\dd u\,\rho_{\rm L}(u)\log(1+\ee^{-\zeta(u)})
  +
  \int\dd u\dd v\,\rho_{\rm p}(v)K^{\rm dens}(v,u)
       \log(1+\ee^{-\zeta(u)}),
\]
which reduces by \eqref{eq:one-species-density-equation} to
\eqref{eq:one-species-log-z}.  The ground-state formula
\eqref{eq:one-species-ground-state-energy} is the usual mirror
space-time reinterpretation of \(\log Z=-\widetilde L E_0(L)\).
\end{proof}

The kernel sign in \eqref{eq:one-species-tba} is the stringbook convention.
If the scattering factor obeys the diagonal unitarity relation
\(S(v,u)S(u,v)=1\), then
\[
  K^{\rm dens}(u,v)
  =
  \frac{1}{2\pi\ii}\partial_v\log S(u,v)
  =
  -
  \frac{1}{2\pi\ii}\partial_v\log S(v,u).
\]
Thus the source-derivative kernel \(K(v,u)\) of
Definition~\ref{def:planar-n4-mirror-kernel-datum} enters the same TBA equation
with a minus sign.  This is the convention used in
\eqref{eq:planar-n4-mirror-tba-general}; the displayed chemical potential
\(\mu_A\) there is the negative of \(\widetilde\mu_A\) in the one-species
statistical-mechanical derivation.

Excited states add one more honest hypothesis: the contour deformation must be
specified.  In the one-species model, inserting moving defects at rapidities
\(u_j\) changes the pseudoenergy to
\[
  \zeta(u)
  =
  \widetilde\beta\,\widetilde\epsilon(u)+\widetilde\mu
  -
  \sum_j\log S(u,u_j)
  +
  \int\dd v\,K^{\rm dens}(u,v)\log(1+\ee^{-\zeta(v)}),
\]
and the defect rapidities are fixed by
\[
  \zeta(u_j)=-2\pi\ii\left(n_j+\frac12\right),
  \qquad n_j\in\mathbb Z.
\]
Equivalently \(1+\ee^{-\zeta}\) has a zero at each defect.  In the planar
\(\mathcal N=4\) problem this condition becomes the exact Bethe-root
regularity condition after analytic continuation to the physical sheet.

\begin{proposition}[Excited-state contour deformation in one species]
\label{prop:one-species-excited-contour-deformation}
Let \(S(u,v)\) be a meromorphic two-particle mirror scattering factor with
no zero or pole on the undeformed real \(v\)-contour, and let
\[
  K^{\rm dens}(u,v)=\frac{1}{2\pi\ii}\partial_v\log S(u,v).
\]
Let \(Y(v)=\ee^{-\zeta(v)}\), and suppose that
\[
  F(v)=1+Y(v)
\]
has simple zeros \(u_1,\ldots,u_K\) crossed by the excited-state contour
deformation and no other singularities are crossed.  Assume also that the
boundary terms at infinity vanish in the integrations by parts below.  With
the contour orientation for which the crossed zeros are enclosed
counterclockwise, the deformed TBA source is
\begin{equation}
  \int_{\mathcal C_{\rm exc}}\dd v\,
  K^{\rm dens}(u,v)\log F(v)
  =
  \int_{\mathbb R}\dd v\,
  K^{\rm dens}(u,v)\log F(v)
  -
  \sum_{j=1}^K \log S(u,u_j).
  \label{eq:one-species-excited-source-residue}
\end{equation}
Moreover, the mirror energy integral obeys
\begin{equation}
  -\int_{\mathcal C_{\rm exc}}\frac{\dd v}{2\pi}\,
  \frac{\dd\widetilde p(v)}{\dd v}\log F(v)
  =
  \sum_{j=1}^K \ii\,\widetilde p(u_j)
  -
  \int_{\mathbb R}\frac{\dd v}{2\pi}\,
  \frac{\dd\widetilde p(v)}{\dd v}\log F(v).
  \label{eq:one-species-excited-energy-residue}
\end{equation}
The root condition is
\begin{equation}
  F(u_j)=0
  \quad\Longleftrightarrow\quad
  \zeta(u_j)=-2\pi\ii\left(n_j+\frac12\right),
  \qquad n_j\in\mathbb Z.
  \label{eq:one-species-excited-root-quantization}
\end{equation}
\end{proposition}

\begin{proof}
For the source term, write the contour difference as the sum of small
positively oriented loops around the crossed zeros and integrate by parts:
\[
  \oint
  \frac{\dd v}{2\pi\ii}\,
  \partial_v\log S(u,v)\log F(v)
  =
  -
  \oint
  \frac{\dd v}{2\pi\ii}\,
  \log S(u,v)\partial_v\log F(v).
\]
Since each \(u_j\) is a simple zero of \(F\),
\(\partial_v\log F(v)\) has residue \(1\) at \(v=u_j\).  The residue theorem
therefore gives the contribution
\(-\sum_j\log S(u,u_j)\), proving
\eqref{eq:one-species-excited-source-residue}.  The energy term is the same
calculation with \(\log S(u,v)\) replaced by \(\ii\widetilde p(v)\):
\[
  -\oint\frac{\dd v}{2\pi}\,
  \widetilde p'(v)\log F(v)
  =
  \oint\frac{\dd v}{2\pi}\,
  \widetilde p(v)\partial_v\log F(v)
  =
  \sum_j \ii\,\widetilde p(u_j).
\]
Finally, \(F(u_j)=0\) means \(\ee^{-\zeta(u_j)}=-1\).  Choosing a logarithm
branch gives \(-\zeta(u_j)=\ii\pi(2n_j+1)\), which is
\eqref{eq:one-species-excited-root-quantization}.
\end{proof}

This proposition fixes the signs in the one-species formula above.  The
minus sign in the driving term is not an optional convention: it is the
integration-by-parts sign in the residue calculation.  The plus sign in the
discrete energy contribution is equally fixed.  After the inverse mirror
continuation of a crossed level-I root, the planar convention is
\(\ii\widetilde p(u_j)=E(u_j)\).  Thus the deformed mirror free-energy
formula produces the physical-magnon energy term in
\eqref{eq:planar-n4-excited-tba-energy}, while the zero condition becomes
the exact Bethe equation on the physical sheet.

Let \(A,B\) range over the mirror Bethe-string species: physical
\(Q\)-particle bound states, auxiliary \(y\)-roots, and auxiliary \(w\)-string
families for the two \(\mathfrak{su}(2|2)\) wings.  Denote the mirror
pseudoenergy by \(\epsilon_A(u)\) and set
\[
  Y_A(u)=\ee^{-\epsilon_A(u)}.
\]
The TBA equations have the form
\begin{equation}
  \epsilon_A(u)
  =
  L\,\widetilde E_A(u)-\mu_A
  -
  \sum_B\int_{\mathcal C_B}\dd v\,
  K_{BA}(v,u)\log\!\left(1+Y_B(v)\right),
  \label{eq:planar-n4-mirror-tba-general}
\end{equation}
with kernels
\[
  K_{BA}(v,u)
  =
  \frac{1}{2\pi\ii}\frac{\partial}{\partial v}
  \log S_{BA}^{\rm mirror}(v,u).
\]
The contours \(\mathcal C_B\), signs for fermionic auxiliary roots, and
chemical potentials \(\mu_A\) are part of the mirror-model definition.  They
are not decorations: changing them changes the finite-size answer.

\begin{proposition}[Mirror-TBA node and source inventory]
\label{prop:planar-n4-mirror-tba-node-source-inventory}
Assume the mirror Bethe-string hypothesis of
Proposition~\ref{prop:planar-n4-mirror-auxiliary-strings}, diagonal scalar
factors for the fused mirror strings satisfying
\[
  S^{\rm sb}_{AB}(u,v)S^{\rm mirror}_{BA}(v,u)=1
\]
on the chosen contours, and no contour-crossing source in the ground-state
calculation.  Let
\[
  \mathcal I
  =
  \{\bullet_Q:Q\geq1\}
  \cup
  \{y_\pm^{(\alpha)},(v|M)^{(\alpha)},(w|M)^{(\alpha)}:
       M\geq1,\ \alpha=L,R\}.
\]
The stringbook symbols are identified as
\[
  y_+^{(\alpha)}=\oplus^{(\alpha)},\qquad
  y_-^{(\alpha)}=\ominus^{(\alpha)},\qquad
  (v|M)^{(\alpha)}=M|yw^{(\alpha)},\qquad
  (w|M)^{(\alpha)}=M|w^{(\alpha)}.
\]
Put \(\nu_L=+1\), \(\nu_R=-1\).  The driving terms are
\[
  d_{\bullet_Q}=L\,\widetilde E_Q,\qquad
  d_{y_\pm^{(\alpha)}}=\nu_\alpha\pi\ii,\qquad
  d_{(v|M)^{(\alpha)}}=d_{(w|M)^{(\alpha)}}=0,
\]
or equivalently \(d_A=L\widetilde E_A-\mu_A\), with
\(\widetilde E_A=0\) for auxiliary nodes and
\(\mu_{y_\pm^{(\alpha)}}=-\nu_\alpha\pi\ii\).  If
\[
  K^{\rm sb}_{AB}(u,v)
  =
  \frac{1}{2\pi\ii}\partial_v\log S^{\rm sb}_{AB}(u,v)
\]
denotes the target-first stringbook kernel, then
\begin{equation}
  K^{\rm sb}_{AB}(u,v)=-K_{BA}(v,u).
  \label{eq:planar-n4-stringbook-source-kernel-bridge}
\end{equation}
Consequently \eqref{eq:planar-n4-mirror-tba-general} is equivalent to the
target-first system
\begin{equation}
  \epsilon_A(u)
  =
  d_A(u)
  +
  \sum_{B\in\mathcal I}
  \int_{\mathcal C_B}\dd v\,
  K^{\rm sb}_{AB}(u,v)\log(1+Y_B(v)).
  \label{eq:planar-n4-mirror-tba-target-first}
\end{equation}
Write \(\mathcal S_A\) for the convolution part of
\eqref{eq:planar-n4-mirror-tba-target-first}.  The nonzero source families
are precisely
\begin{align*}
  \mathcal S_{\bullet_Q}
  &=
  \sum_{Q'\geq1}K^{\bullet\bullet}_{QQ'}*\log(1+Y_{\bullet_{Q'}})
  +
  \sum_{\alpha=L,R}\sum_{M\geq2}
    K^{\bullet v}_{QM}*\log(1+Y_{(v|M)^{(\alpha)}})
  \\
  &\quad
  +
  \sum_{\alpha=L,R}
  \left[
    K^{\bullet y_+}_{Q}*\log(1+Y_{y_+^{(\alpha)}})
    +
    K^{\bullet y_-}_{Q}*\log(1+Y_{y_-^{(\alpha)}})
  \right]
  ,
  \\
  \mathcal S_{y_\pm^{(\alpha)}}
  &=
  \sum_{Q\geq1}K^{y_\pm\bullet}_{Q}*\log(1+Y_{\bullet_Q})
  +
  \sum_{M\geq1}K_M*
  \log\frac{1+Y_{(v|M)^{(\alpha)}}}
           {1+Y_{(w|M)^{(\alpha)}}}
  ,
  \\
  \mathcal S_{(v|M)^{(\alpha)}}
  &=
  \sum_{Q\geq1}K^{v\bullet}_{MQ}*\log(1+Y_{\bullet_Q})
  +
  \sum_{N\geq1}K_{MN}*\log(1+Y_{(v|N)^{(\alpha)}})
  \\
  &\quad
  +
  K_M*\log\frac{1+Y_{y_+^{(\alpha)}}}
                 {1+Y_{y_-^{(\alpha)}}}
  ,
  \\
  \mathcal S_{(w|M)^{(\alpha)}}
  &=
  \sum_{N\geq1}K_{MN}*\log(1+Y_{(w|N)^{(\alpha)}})
  +
  K_M*\log\frac{1+Y_{y_+^{(\alpha)}}}
                 {1+Y_{y_-^{(\alpha)}}}.
\end{align*}
The lower bound \(M\geq2\) in the \(\bullet_Q\) equation is the stringbook
indexing of the \(v\)-string source adjacent to the \(y\)-sheet boundary.
The denominator entries are not optional signs: they encode the reversed
ordering of the \(\ominus\)-sheet and \(w\)-string Bethe roots.
\end{proposition}

\begin{proof}
The one-species derivation above already gives the density-to-TBA transpose.
For species \(A\), the logarithmic Bethe-Yang equation is differentiated in
the target rapidity.  The entropy variation transposes the density kernel, so
the target-first TBA contains
\((2\pi\ii)^{-1}\partial_v\log S^{\rm sb}_{AB}(u,v)\).  The diagonal
unitarity relation displayed in the hypotheses gives
\[
  \frac{1}{2\pi\ii}\partial_v\log S^{\rm sb}_{AB}(u,v)
  =
  -
  \frac{1}{2\pi\ii}\partial_v\log S^{\rm mirror}_{BA}(v,u),
\]
which is \eqref{eq:planar-n4-stringbook-source-kernel-bridge} and converts
\eqref{eq:planar-n4-mirror-tba-target-first} to
\eqref{eq:planar-n4-mirror-tba-general}.  Only the \(\bullet_Q\) strings carry
mirror energy, so the length term is \(L\widetilde E_Q\) on those nodes and is
zero on auxiliary nodes.  The sign \(-1\) in the level-II mirror Bethe-Yang
equation gives the additive fermion chemical potential
\(\nu_\alpha\pi\ii\) in each wing, equivalently
\(\mu_{y_\pm^{(\alpha)}}=-\nu_\alpha\pi\ii\) in the source-kernel convention.

The remaining entries are the fused forms of
\eqref{eq:planar-n4-mirror-auxiliary-bethe-yang}.  Fusing level-I roots gives
the \(\bullet\)-kernels; fusing the auxiliary arrays gives the rational
\(K_M\) and \(K_{MN}\) kernels used below.  The roots on the
\(\ominus\)-sheet and in the \(w\)-strings are ordered oppositely to the
\(\oplus\)-sheet and \(v\)-string roots in the stringbook Bethe-Yang
products.  Their scattering factors therefore enter inverted, which is
exactly the logarithmic ratio displayed in the \(y\)-, \(v\)-, and
\(w\)-node equations.
\end{proof}

\begin{proposition}[Fused mirror kernels in stringbook orientation]
\label{prop:planar-n4-fused-mirror-kernel-formulas}
Fix a mirror sheet and write
\[
  \mathcal R_m=
  \left\{-\frac{m-1}{2},-\frac{m-1}{2}+1,\ldots,\frac{m-1}{2}\right\}.
\]
For an integer \(a\), \(x^{[a]}(u)\) denotes the continuation of
\(x(u+\ii a/2)\) on that sheet, with the side of the cut fixed by the
contour datum.  Away from zeros and poles, the fused rational
\(\bullet_m\)-\(\bullet_n\) phase is
\begin{align}
  \Phi^{\bullet\bullet}_{mn}(u,v)
  &=
  \sum_{r\in\mathcal R_m}
  \sum_{s\in\mathcal R_n}
  \log\left[
    \frac{x^{[2s-1]}(v)-x^{[2r+1]}(u)}
         {x^{[2s+1]}(v)-x^{[2r-1]}(u)}
    \frac{1-\bigl(x^{[2s+1]}(v)x^{[2r-1]}(u)\bigr)^{-1}}
         {1-\bigl(x^{[2s-1]}(v)x^{[2r+1]}(u)\bigr)^{-1}}
  \right],
  \label{eq:planar-n4-mirror-bullet-bullet-rational-phase}
\end{align}
and the dressing contribution is the endpoint combination
\begin{align}
  \Theta^{\bullet\bullet}_{mn}(u,v)
  =
  2\ii\bigl(&
  \chi(x^{[m]}(u),x^{[n]}(v))
  +
  \chi(x^{[-m]}(u),x^{[-n]}(v))
  \nonumber\\
  &-
  \chi(x^{[-m]}(u),x^{[n]}(v))
  -
  \chi(x^{[m]}(u),x^{[-n]}(v))
  \bigr).
  \label{eq:planar-n4-mirror-bullet-bullet-dressing-phase}
\end{align}
Thus
\[
  K^{\bullet\bullet}_{mn}(u,v)
  =
  \frac{1}{2\pi\ii}\partial_v
  \left(
    \Phi^{\bullet\bullet}_{mn}(u,v)
    +
    \Theta^{\bullet\bullet}_{mn}(u,v)
  \right)
\]
in the target-first convention.  The \(y\)-sheet and \(v\)-string phases
that occur in the source inventory are
\begin{align}
  \log S^{y_+\bullet}_n(u,v)
  &=
  \log\left[
    \frac{x(u)-x^{[n]}(v)}
         {x(u)-x^{[-n]}(v)}
    \sqrt{\frac{x^{[-n]}(v)}{x^{[n]}(v)}}
  \right],
  \label{eq:planar-n4-mirror-yplus-bullet-phase}\\
  \log S^{y_-\bullet}_n(u,v)
  &=
  \log\left[
    \frac{x(u)^{-1}-x^{[n]}(v)}
         {x(u)^{-1}-x^{[-n]}(v)}
    \sqrt{\frac{x^{[-n]}(v)}{x^{[n]}(v)}}
  \right],
  \label{eq:planar-n4-mirror-yminus-bullet-phase}\\
  \log S^{(v|m)\bullet}_{mn}(u,v)
  &=
  \sum_{\sigma=\pm1}
  \sum_{r\in\mathcal R_m}
  \sum_{s\in\mathcal R_n}
  \log\left[
    \frac{x^{[2r+\sigma]}(u)-x^{[2s+1]}(v)}
         {x^{[2r+\sigma]}(u)-x^{[2s-1]}(v)}
    \sqrt{\frac{x^{[2s-1]}(v)}{x^{[2s+1]}(v)}}
  \right].
  \label{eq:planar-n4-mirror-v-bullet-phase}
\end{align}
Their opposite orientations are reciprocals:
\[
  \log S^{\bullet y_\pm}_n(v,u)=-\log S^{y_\pm\bullet}_n(u,v),
  \qquad
  \log S^{\bullet(v|m)}_{nm}(v,u)
  =
  -\log S^{(v|m)\bullet}_{mn}(u,v).
\]
Finally, the auxiliary rational string phase
\[
  \Phi^{\rm aux}_{mn}(u,v)
  =
  \sum_{r\in\mathcal R_m}\sum_{s\in\mathcal R_n}
  \log\frac{u-v+\ii(r-s+1)}{u-v+\ii(r-s-1)}
\]
has kernel
\begin{equation}
  \frac{1}{2\pi\ii}\partial_v\Phi^{\rm aux}_{mn}
  =
  K_{m+n}+K_{\lvert m-n\rvert}
  +
  2\sum_{j=1}^{\min(m,n)-1}K_{\lvert m-n\rvert+2j},
  \label{eq:planar-n4-mirror-auxiliary-fused-kernel-crosswalk}
\end{equation}
with \(K_0=0\) and
\[
  K_N(u-v)
  =
  \frac{1}{2\pi\ii}\partial_v
  \log\frac{u-v+\ii N/2}{u-v-\ii N/2}.
\]
\end{proposition}

\begin{proof}
The rational factors are finite products over the constituents of the two
Bethe strings.  The sets \(\mathcal R_m\) and \(\mathcal R_n\) are exactly the
constituent shifts, so multiplying the elementary level-I rational factor
over \(r\) and \(s\) gives
\eqref{eq:planar-n4-mirror-bullet-bullet-rational-phase}.  Reversing the
order of the two strings in each elementary factor inverts the product, which
is the reciprocal orientation stated for \(y_\pm\) and \((v|m)\) sources.
The distinction between \(y_+\) and \(y_-\) is the replacement
\(x(u)\mapsto x(u)^{-1}\) on the two \(y\)-sheets.

For the purely auxiliary kernel put \(z=u-v\).  The derivative of
\(\log(z+\ii a)\) with respect to \(v\) is \(-(z+\ii a)^{-1}\).  Hence the
double sum contributes the difference between the multiplicities of
\(r-s-1\) and \(r-s+1\).  As \(r\) and \(s\) run over two consecutive
arithmetic strings, all interior multiplicities cancel in pairs except the
two boundary terms and the doubled interior terms with the same parity as
\(|m-n|\).  This gives precisely the right-hand side of
\eqref{eq:planar-n4-mirror-auxiliary-fused-kernel-crosswalk}; it is the fused
\(A_\infty\) kernel used in the Y-system inversion below.

The dressing term is different in status.  If \(\chi(x,y)=-\chi(y,x)\) on the
continued branches, the endpoint combination
\eqref{eq:planar-n4-mirror-bullet-bullet-dressing-phase} is antisymmetric
under \((m,u)\leftrightarrow(n,v)\), so it is compatible with scalar
unitarity.  The existence of the globally continued \(\chi\), its branch
choice, and possible additional crossed singularities are part of the
dressing-phase analytic-continuation input inherited from
Chapter~\ref{ch:planar-n4-asymptotic-bethe-ansatz}.  The finite fusion algebra
proved here supplies the rational products and the endpoint sign package.
\end{proof}

For a state with physical Bethe roots \(u_1,\ldots,u_K\), the excited-state
TBA is obtained by contour deformation.  The physical roots satisfy exact
Bethe equations of the form
\[
  1+Y_{\bullet_1}^{\rm phys}(u_j)=0,
  \qquad j=1,\ldots,K,
\]
and the energy is
\begin{equation}
  \Delta-J
  =
  \sum_{j=1}^K E(u_j)
  -
  \sum_{Q=1}^\infty
  \int_{\R}\frac{\dd u}{2\pi}\,
  \frac{\dd\widetilde p_Q}{\dd u}
  \log(1+Y_{\bullet_Q}(u)).
  \label{eq:planar-n4-excited-tba-energy}
\end{equation}
The first term is the asymptotic magnon energy.  The integral term is the
finite-size wrapping correction from the thermal mirror bath.

\section{Y-System}

The passage from TBA integral equations to local Y-system relations is not a
formal cancellation of symbols.  It uses a special inverse for the fused
Bethe-string kernels and an analyticity strip in which
\(\mathsf s^{-1}\) may be represented by shifts.  We spell out the
auxiliary-wing case because it is the cleanest part of the planar
\(\mathcal N=4\) construction and fixes the normalization used later by the
T-hook notation.

\begin{proposition}[Auxiliary-string kernel inverse]
\label{prop:planar-n4-auxiliary-kernel-inverse}
Let
\[
  (K*f)(u)=\int_{\mathbb R}\dd v\,K(u-v)f(v)
\]
and, for \(n\geq 1\),
\begin{equation}
  K_n(u-v)
  =
  \frac{1}{2\pi\ii}\partial_v
  \log\frac{u-v+\ii n/2}{u-v-\ii n/2}
  =
  \frac{n/2}{\pi((u-v)^2+n^2/4)} .
  \label{eq:planar-n4-rational-string-kernel}
\end{equation}
Set \(K_0=0\), and define the fused \(A_\infty\)-string kernel
\[
  K_{mn}
  =
  K_{m+n}+K_{\lvert m-n\rvert}
  +
  2\sum_{j=1}^{\min(m,n)-1}K_{\lvert m-n\rvert+2j},
  \qquad m,n\geq 1,
\]
with \(A_{mn}=\delta_{mn}\delta+K_{mn}\).  Finally put
\[
  \mathsf s(u-v)=\frac{1}{2\cosh\pi(u-v)} .
\]
On the Fourier side, with
\(\widehat f(\omega)=\int_{\mathbb R}\ee^{\ii\omega u}f(u)\,\dd u\) and
\(q=\ee^{-|\omega|/2}\),
\begin{equation}
  \widehat K_n=q^n,\qquad
  \widehat{\mathsf s}=\frac{q}{1+q^2},\qquad
  \widehat A_{mn}
  =
  \frac{1+q^2}{1-q^2}\left(q^{|m-n|}-q^{m+n}\right).
  \label{eq:planar-n4-a-kernel-fourier-symbol}
\end{equation}
Consequently
\begin{equation}
  (A^{-1})_{mn}
  =
  \delta_{mn}\delta
  -
  (\delta_{m+1,n}+\delta_{m-1,n})\mathsf s,
  \qquad \delta_{0n}=0.
  \label{eq:planar-n4-a-kernel-inverse}
\end{equation}
Equivalently,
\begin{align}
  \sum_{\ell\geq 1}(A^{-1})_{m\ell}*K_{\ell n}
  &=
  (\delta_{m+1,n}+\delta_{m-1,n})\mathsf s,
  \label{eq:planar-n4-a-inverse-kmn}\\
  \sum_{\ell\geq 1}(A^{-1})_{m\ell}*K_\ell
  &=
  \delta_{m1}\mathsf s .
  \label{eq:planar-n4-a-inverse-kn}
\end{align}
\end{proposition}

\begin{proof}
The transform of \eqref{eq:planar-n4-rational-string-kernel} is the elementary
Poisson-kernel identity \(\widehat K_n(\omega)=\ee^{-n|\omega|/2}\).  The
displayed expression for \(\widehat A_{mn}\) follows by summing the finite
geometric progression in the fused kernel.  In Fourier variables the asserted
inverse is the tridiagonal matrix
\[
  \delta_{mn}
  -
  \frac{q}{1+q^2}(\delta_{m+1,n}+\delta_{m-1,n}).
\]
Multiplying it by the last expression in
\eqref{eq:planar-n4-a-kernel-fourier-symbol} gives
\[
  \widehat A_{mn}
  -
  \widehat{\mathsf s}\,
  (\widehat A_{m+1,n}+\widehat A_{m-1,n})
  =
  \delta_{mn},
  \qquad \widehat A_{0n}=0.
\]
For \(m\neq n\) the two geometric tails cancel.  For \(m=n\) the remaining
jump is one.  Equations \eqref{eq:planar-n4-a-inverse-kmn} and
\eqref{eq:planar-n4-a-inverse-kn} follow by subtracting the identity kernel
from \(A\) and by applying the same calculation to the one-index kernel
\(\widehat K_n=q^n\).
\end{proof}

\begin{proposition}[One auxiliary wing of the Y-system]
\label{prop:planar-n4-auxiliary-wing-y-system}
Fix one \(\mathfrak{su}(2|2)\) wing.  Let \(Y_n^\circ\) denote the \(n|w\)
string Y-functions, and let \(Y^\oplus,Y^\ominus\) denote the two \(y\)-sheet
Y-functions in the stringbook orientation.  Suppose the \(n|w\) part of the
mirror TBA can be written as
\begin{equation}
  \sum_{m\geq 1}A_{nm}*\log Y_m^\circ
  =
  \sum_{m\geq 1}K_{nm}*\log(1+Y_m^\circ)
  +
  K_n*\log\frac{1+1/Y^\oplus}{1+1/Y^\ominus},
  \label{eq:planar-n4-wing-tba-a-form}
\end{equation}
with \(Y_0^\circ=0\).  Assume all functions appearing after convolution by
\(\mathsf s\) are analytic in the strip \(|\operatorname{Im}u|<1/2\), decay
enough at infinity, and have no pole crossed by the shifts
\(u\mapsto u\pm\ii/2\).  Then
\begin{equation}
  \frac{(Y_n^\circ)^+(Y_n^\circ)^-}
       {(1+Y_{n+1}^\circ)(1+Y_{n-1}^\circ)}
  =
  \left(
    \frac{1+1/Y^\oplus}{1+1/Y^\ominus}
  \right)^{\delta_{n1}},
  \qquad n\geq 1.
  \label{eq:planar-n4-wing-y-system}
\end{equation}
\end{proposition}

\begin{proof}
Apply \(A^{-1}\) to \eqref{eq:planar-n4-wing-tba-a-form}.  Using
\eqref{eq:planar-n4-a-inverse-kmn} and \eqref{eq:planar-n4-a-inverse-kn}
gives
\[
  \log Y_n^\circ
  =
  \mathsf s*\log(1+Y_{n+1}^\circ)
  +
  \mathsf s*\log(1+Y_{n-1}^\circ)
  +
  \delta_{n1}\mathsf s*
  \log\frac{1+1/Y^\oplus}{1+1/Y^\ominus}.
\]
For functions analytic in the stated strip,
\(\mathsf s^{-1}*f=f^+ + f^-\).  Applying this inverse and exponentiating gives
\eqref{eq:planar-n4-wing-y-system}.  If a pole or branch cut crosses during
the shift, the right-hand side acquires extra source terms; this is precisely
why the analytic data below are part of the finite-volume spectral problem.
\end{proof}

\begin{proposition}[Data lost by the \(\mathsf s\)-kernel inverse]
\label{prop:planar-n4-s-inverse-data-loss}
Write
\[
  (\nabla_{\mathsf s}F)(u)=F(u+\ii/2)+F(u-\ii/2).
\]
Suppose \(F=\mathsf s*f\), where \(f\) is integrable, \(F\) is holomorphic in
a neighbourhood of the closed strip \(|\operatorname{Im}u|\leq1/2\), and the
two contour shifts meet no singularity of \(F\).  In logarithmic TBA
applications this means that no zero or pole of the underlying Y-function, and
no logarithmic cut, is crossed.  Then
\begin{equation}
  \nabla_{\mathsf s}F=f .
  \label{eq:planar-n4-s-inverse-regular-domain}
\end{equation}
The converse datum is not contained in the local equation.  For
\(\alpha\in\C\),
\begin{equation}
  Z_\alpha(u)=\frac{1}{2\cosh\pi(u-\alpha)}
  \label{eq:planar-n4-s-inverse-zero-mode}
\end{equation}
satisfies
\[
  Z_\alpha(u+\ii/2)+Z_\alpha(u-\ii/2)=0
\]
as a meromorphic identity away from \(u=\alpha\), while its closest poles are
at \(u=\alpha\pm\ii/2\).  Thus applying \(\mathsf s^{-1}\) determines only the
regular part of a logarithmic TBA equation modulo such boundary data.  If the
logarithm instead contains the shifted zero-pole factor
\[
  B_\alpha(u)=\frac{u-\alpha+\ii/2}{u-\alpha-\ii/2},
\]
then
\begin{equation}
  B_\alpha^+(u)B_\alpha^-(u)
  =
  \frac{u-\alpha+\ii}{u-\alpha-\ii}.
  \label{eq:planar-n4-s-inverse-source-memory}
\end{equation}
Consequently the local Y-system knows neither the positions \(\alpha\), nor
the integer powers of such factors, nor the side of the cut from which they
come.  These are extra analytic data supplied by the mirror-TBA contour and
by exact-root regularity.
\end{proposition}

\begin{proof}
With the Fourier convention of
Proposition~\ref{prop:planar-n4-auxiliary-kernel-inverse},
\(\widehat{\mathsf s}(\omega)=1/(2\cosh(\omega/2))\).  Shifting the contour
for \(F(u+\ii/2)\) and \(F(u-\ii/2)\) back to the real line multiplies the
Fourier transform by \(\ee^{\omega/2}\) and \(\ee^{-\omega/2}\), respectively.
Under the stated no-crossing hypothesis the sum therefore has transform
\[
  2\cosh(\omega/2)\,\widehat{\mathsf s}(\omega)\widehat f(\omega)
  =
  \widehat f(\omega),
\]
which proves \eqref{eq:planar-n4-s-inverse-regular-domain}.  If a singularity
lies on a shifted contour, the same contour move has a residue or jump
contribution; that contribution is not present in the pointwise local
equation unless it is recorded as a source.

For the zero mode, use
\[
  \cosh\pi(u-\alpha+\ii/2)=\ii\sinh\pi(u-\alpha),
  \qquad
  \cosh\pi(u-\alpha-\ii/2)=-\ii\sinh\pi(u-\alpha).
\]
The two shifted terms cancel away from their boundary pole.  The last identity
is the direct multiplication of \(B_\alpha(u+\ii/2)\) and
\(B_\alpha(u-\ii/2)\).
\end{proof}

The TBA equations imply functional relations among Y-functions after applying
kernel identities and shifting the spectral parameter.  The bulk of the
planar \(\mathcal N=4\) T-hook is the subset
\[
  \mathfrak T_{\rm bulk}
  =
  \{(a,s): a\in\mathbb Z_{\ge1},\ s\in\mathbb Z,\ |s|\le2\}
  \cup
  \{(a,s): a=1,2,\ s\in\mathbb Z\}.
\]
The row \(a=0\) is adjoined as a gauge boundary \(T_{0,s}\); outside
\(\mathfrak T_{\rm bulk}\) and this boundary row one sets \(T_{a,s}=0\).  A
Y-node is called interior when \(a\geq1\) and all four neighbouring
T-functions appearing below belong to \(\mathfrak T_{\rm bulk}\) or to the
boundary row, and are non-zero on the patch under consideration.  On such
nodes the local relation has the form
\begin{equation}
  Y_{a,s}^+(u)Y_{a,s}^-(u)
  =
  \frac{(1+Y_{a,s+1}(u))(1+Y_{a,s-1}(u))}
       {(1+Y_{a+1,s}(u)^{-1})(1+Y_{a-1,s}(u)^{-1})},
  \label{eq:planar-n4-y-system}
\end{equation}
where
\[
  f^\pm(u)=f(u\pm\ii/2).
\]
Boundary nodes require modified relations because one of the four neighboring
T-functions has left the hook.

\begin{proposition}[Hirota origin of the interior T-hook Y-system]
\label{prop:planar-n4-hirota-origin-y-system}
Let \(T_{a,s}\) be non-zero on an interior patch of the T-hook and obey
Hirota,
\[
  T_{a,s}^+T_{a,s}^-
  =
  T_{a+1,s}T_{a-1,s}+T_{a,s+1}T_{a,s-1}.
\]
Define
\[
  Y_{a,s}
  =
  \frac{T_{a,s+1}T_{a,s-1}}{T_{a+1,s}T_{a-1,s}} .
\]
Then \(Y_{a,s}\) obeys
\eqref{eq:planar-n4-y-system} at every interior node.  Moreover \(Y_{a,s}\)
is invariant under the local Hirota gauge transformation
\begin{equation}
  T_{a,s}\mapsto
  g_1^{[a+s]}g_2^{[a-s]}g_3^{[-a-s]}g_4^{[-a+s]}T_{a,s},
  \label{eq:planar-n4-t-gauge-redundancy}
\end{equation}
where \(f^{[n]}(u)=f(u+\ii n/2)\).
\end{proposition}

\begin{proof}
Hirota at \((a,s)\) gives the two useful identities
\[
  1+Y_{a,s}
  =
  \frac{T_{a,s}^+T_{a,s}^-}{T_{a+1,s}T_{a-1,s}},
  \qquad
  1+\frac1{Y_{a,s}}
  =
  \frac{T_{a,s}^+T_{a,s}^-}{T_{a,s+1}T_{a,s-1}} .
\]
Now compute
\[
  Y_{a,s}^+Y_{a,s}^-
  =
  \frac{
    T_{a,s+1}^+T_{a,s+1}^-
    T_{a,s-1}^+T_{a,s-1}^-
  }{
    T_{a+1,s}^+T_{a+1,s}^-
    T_{a-1,s}^+T_{a-1,s}^-
  } .
\]
Applying the two displayed identities to the four neighboring nodes turns the
numerator into
\[
  (1+Y_{a,s+1})(1+Y_{a,s-1})
  T_{a+1,s+1}T_{a-1,s+1}T_{a+1,s-1}T_{a-1,s-1},
\]
and the denominator into
\[
  (1+Y_{a+1,s}^{-1})(1+Y_{a-1,s}^{-1})
  T_{a+1,s+1}T_{a+1,s-1}T_{a-1,s+1}T_{a-1,s-1}.
\]
The common T-product cancels, proving
\eqref{eq:planar-n4-y-system}.

For the gauge statement, let
\[
  \mathcal G(a,s)
  =
  \{(1,a+s),(2,a-s),(3,-a-s),(4,-a+s)\}
\]
record which \(g_i\) appears with which shift.  The numerator of \(Y_{a,s}\)
has gauge multiset
\[
  \mathcal G(a,s+1)\sqcup\mathcal G(a,s-1),
\]
and the denominator has
\[
  \mathcal G(a+1,s)\sqcup\mathcal G(a-1,s).
\]
These two multisets are identical.  Hence all gauge factors cancel.  The
calculation is local; analytic gauge choices in the QSC additionally fix
cuts, asymptotics, and gluing data.
\end{proof}

\begin{warning}[Y-system data are not only equations]
\label{warn:y-system-not-complete}
The functional relations \eqref{eq:planar-n4-y-system} do not by themselves
determine a spectrum.  A solution also requires analyticity domains,
large-\(u\) asymptotics, branch-cut discontinuity relations, exact Bethe-root
conditions, and state labels.  Treating the Y-system equations alone as a
definition of the theory discards the information that distinguishes physical
states.
\end{warning}

\begin{proposition}[Local source factor from a shifted zero-pole pair]
\label{prop:planar-n4-local-y-source-factor}
Let \(\alpha\in\C\), and define the elementary shifted source factor
\[
  B_\alpha(u)=\frac{u-\alpha+\ii/2}{u-\alpha-\ii/2}.
\]
Thus \(B_\alpha\) has a simple zero at \(u=\alpha-\ii/2\) and a simple pole at
\(u=\alpha+\ii/2\).  If a meromorphic Y-function is written locally as
\[
  Y_{a,s}(u)=B_\alpha(u)^\nu\,\widehat Y_{a,s}(u),
  \qquad \nu\in\mathbb Z,
\]
where \(\widehat Y_{a,s}\) has no zero or pole crossed by the two shifts
\(u\mapsto u\pm\ii/2\), then
\begin{equation}
  Y_{a,s}^+(u)Y_{a,s}^-(u)
  =
  \left(\frac{u-\alpha+\ii}{u-\alpha-\ii}\right)^\nu
  \widehat Y_{a,s}^+(u)\widehat Y_{a,s}^-(u).
  \label{eq:planar-n4-local-y-source-factor}
\end{equation}
For a finite set of shifted zero-pole pairs, the source factor is the product
of the corresponding powers of
\((u-\alpha+\ii)/(u-\alpha-\ii)\).
\end{proposition}

\begin{proof}
The claim is the algebra of the shift operation.  Directly,
\[
  B_\alpha^+(u)
  =
  \frac{u-\alpha+\ii}{u-\alpha},
  \qquad
  B_\alpha^-(u)
  =
  \frac{u-\alpha}{u-\alpha-\ii}.
\]
Multiplying the two expressions gives
\[
  B_\alpha^+(u)B_\alpha^-(u)
  =
  \frac{u-\alpha+\ii}{u-\alpha-\ii}.
\]
Raising to the integer power \(\nu\) and multiplying by
\(\widehat Y_{a,s}^+\widehat Y_{a,s}^-\) proves
\eqref{eq:planar-n4-local-y-source-factor}.  A finite product follows by
commutativity.
\end{proof}

This proposition is the local algebraic shadow of contour deformation.  The
homogeneous Y-system records what happens after \(\mathsf s^{-1}\) is applied
in a strip free of singularities.  If a zero, pole, or logarithmic branch point
crosses one of the two strip boundaries, the remaining regular part
\(\widehat Y_{a,s}\) still obeys the same shift algebra, but the singular part
contributes the rational source factor in
\eqref{eq:planar-n4-local-y-source-factor}.  The Y-system equations alone do
not determine the positions \(\alpha\), the powers \(\nu\), or the sheet on
which the factor lives; these data come from the excited-state contours and
from the exact Bethe-root regularity conditions.

\section{Analytic Y-System Data}

The analytic Y-system supplements \eqref{eq:planar-n4-y-system} by specifying
where each shifted function is meromorphic and which discontinuities occur
when a shift crosses a cut.  This is not cosmetic.  The symbolic Y-system is a
local relation; the analytic data determine which local solution is the
finite-volume spectrum.

We use the shifted notation
\[
  f^{[n]}(u)=f(u+\ii n/2)
\]
and write
\[
  Z_0^\vee=[-2g,2g]
\]
for the central mirror-sheet Zhukovsky cut.  Let \(\mathcal A_\ell\) denote the
class of functions meromorphic in the open strip
\begin{equation}
  -\frac{\ell}{2}<\operatorname{Im}u<\frac{\ell}{2},
  \label{eq:planar-n4-analytic-strip}
\end{equation}
after removing state-dependent poles and zeros that are explicitly recorded by
source factors.  The basic stringbook-to-T-hook identification is
\begin{align}
  Y^\bullet_n=Y_{n,0},\qquad
  Y^{\oplus,L/R}=Y_{1,\pm1}^{-1},\qquad
  Y^{\ominus,L/R}=Y_{2,\pm2},
  \nonumber\\
  Y^{\vartriangle,L/R}_n=Y_{n+1,\pm1}^{-1},\qquad
  Y^{\circ,L/R}_n=Y_{1,\pm(n+1)}^{-1}.
  \label{eq:planar-n4-y-node-stringbook-map}
\end{align}

\begin{assumption}[Analytic Y-system strip and cut data]
\label{ass:planar-n4-analytic-y-system-data}
Fix the mirror-TBA sheet and the stringbook node convention
\eqref{eq:planar-n4-y-node-stringbook-map}.  For every \(n\geq1\),
\begin{equation}
  Y_{n,0},\qquad Y_{n+1,\pm1},\qquad Y_{1,\pm(n+1)}
  \in \mathcal A_n .
  \label{eq:planar-n4-y-strip-assumption}
\end{equation}
Their first Zhukovsky branch points occur at
\begin{equation}
  u=\pm 2g+\frac{\ii\ell}{2},
  \qquad
  \ell=\pm n,\ \pm(n+2),\ \pm(n+4),\ldots .
  \label{eq:planar-n4-y-branch-lattice}
\end{equation}
The central fermionic nodes \(Y_{1,\pm1}\) and \(Y_{2,\pm2}\) have cuts
\(Z_0^\vee+\ii m\), \(m\in\mathbb Z\), and their two sides on the central cut
are related by
\begin{equation}
  Y_{2,\pm2}(u+\ii0)=\frac{1}{Y_{1,\pm1}(u-\ii0)},
  \qquad u\in Z_0^\vee .
  \label{eq:planar-n4-central-fermion-cut-inversion}
\end{equation}
Finally, every zero or pole introduced by an excited-state contour deformation
is part of the data and must be recorded by a finite product of local factors
of the form in Proposition~\ref{prop:planar-n4-local-y-source-factor}.
\end{assumption}

The strip statement is an analytic input extracted from the mirror TBA and its
asymptotic Bethe-ansatz limit; it is not a theorem of the local Hirota algebra.
The same limitation is visible in the Arutyunov--Frolov derivation of the
mirror TBA: applying the inverse \(s\)-kernel loses data, and the local
Y-equations are controlled on the central mirror interval unless additional
analytic-continuation assumptions are supplied.  We do not use that literature
statement as a proof.  The assumption above is the independent monograph datum
that must be checked against the chosen mirror sheet, source factors, and
dressing branch.
The branch-point lattice explains why the homogeneous Y-system can be shifted
inside the strip \(\mathcal A_n\) but acquires sources when an endpoint, pole,
zero, or dressing cut is crossed.  The central inversion
\eqref{eq:planar-n4-central-fermion-cut-inversion} is the piece that later
turns into the magic-sheet gluing condition for the one-row T-gauge and,
eventually, into the \(P\mu\)-system.

With these conventions the exact Bethe equations are regularity conditions:
\[
  1+Y_{\bullet_1}^{\rm phys}(u_j)=0
\]
removes would-be singularities at the physical roots after contour
deformation.  This formulation is deliberately more verbose than the symbolic
Y-system.  The shifts \(u\mapsto u\pm\ii/2\) can move a point from the physical
sheet to the mirror sheet, and the operations ``shift'' and ``continue through
a cut'' do not commute without recording the path.

\section{Konishi as the First Wrapping Test}

The Konishi multiplet contains an \(SL(2)\) descendant represented by a
length-two state with two derivative magnons.  In the asymptotic Bethe ansatz,
the two physical roots are opposite,
\[
  u_1=-u_2.
\]
At weak coupling the one-loop solution is
\[
  u_1=\frac{1}{2\sqrt3},\qquad
  u_2=-\frac{1}{2\sqrt3},
\]
in the normalization of the rational Bethe equations.  The asymptotic
dimension matches direct perturbative gauge theory through the orders before
wrapping appears.  The first wrapping correction is supplied by the mirror
integral term in \eqref{eq:planar-n4-excited-tba-energy}.  Conceptually, this
is the finite-length correction to the trace, not a correction to the
two-magnon dispersion relation.

The exact finite-coupling Konishi dimension is obtained by solving the
excited-state TBA or, more efficiently, the quantum spectral curve of the next
chapter.  In this monograph the TBA formula is used as a logically explicit
construction of the finite-size correction: every input is an S-matrix kernel,
a contour, a chemical potential, or a Bethe-root condition.

At weak coupling the leading finite-size effect is the Luescher or first
wrapping correction.  The \(SL(2)\) Konishi descendant has \(K=2\) physical
roots \(u_1=-u_2\) with
\[
  u_1=\frac{1}{2\sqrt3}
  \bigl(1+8g^2-20g^4+(112+48\zeta(3))g^6+O(g^8)\bigr)
\]
in the asymptotic equation.  The asymptotic energy gives
\[
  \Delta_{\rm K}^{\rm ABA}
  =
  4+12g^2-48g^4+336g^6-(2820+288\zeta(3))g^8+O(g^{10}).
\]

\begin{assumption}[Leading Konishi mirror input]
\label{ass:konishi-leading-mirror-input}
The leading weak-coupling wrapping computation uses the following mirror-TBA
data.
First, the physical roots in all factors contributing at order \(g^8\) are
the one-loop values
\[
  u_\ast=\frac{1}{2\sqrt3},\qquad u_1=u_\ast,\qquad u_2=-u_\ast .
\]
Second, the virtual particles are the mirror \(Q\)-particle bound states,
\(Q\in\mathbb Z_{>0}\), in the antisymmetric \(SL(2)\)-type mirror sector.
Third, for real mirror rapidity \(u\) the weak mirror branch is
\[
  x^{[Q]}(u)=\frac{g}{u+\ii Q/2}+O(g^3),\qquad
  x^{[\ell]}(u)=\frac{u+\ii\ell/2}{g}+O(g),
  \qquad -Q\leq \ell\leq Q-1.
\]
Fourth, the leading contribution of the dressing phase to the asymptotic
mirror \(Y_Q\)-function starts beyond the displayed \(g^8\) coefficient.
Finally, the excited-state contour deformation is the one for which
\eqref{eq:planar-n4-excited-tba-energy} has no additional weak-coupling
\(\mu\)-term for Konishi.
\end{assumption}

Under this assumption the asymptotic mirror density has the stringbook form
\begin{equation}
  Y_{\bullet_Q}(u)
  =
  g^8\,Y_Q^{(0)}(u)+O(g^{10}),
  \label{eq:konishi-leading-yq}
\end{equation}
where
\begin{equation}
\begin{aligned}
  Y_Q^{(0)}(u)
  &=
  \frac{
    4Q^2\left(u^2-u_\ast^2+(Q^2-1)/4\right)^2
  }{
    \left(u^2+Q^2/4\right)^4
  }
\\
  &\quad\times
  \prod_{\varepsilon=\pm1}
  \frac{1}{
    \left((u-\varepsilon u_\ast)^2+(Q+1)^2/4\right)
    \left((u-\varepsilon u_\ast)^2+(Q-1)^2/4\right)
  } .
\end{aligned}
\label{eq:konishi-yq0}
\end{equation}
All symbols in \eqref{eq:konishi-yq0} are now fixed: \(Q\) is the positive
integer mirror bound-state charge, \(u\in\mathbb R\) is the mirror rapidity,
and \(g=\sqrt\lambda/(4\pi)\) is the coupling of
Chapter~\ref{ch:planar-n4-asymptotic-bethe-ansatz}.

The mirror momentum measure is also a branch-sensitive datum.  In the same
convention as the stringbook, the mirror \(Q\)-particle satisfies
\[
  \ii\widetilde p_Q(u)
  =
  -Q-2\ii g\bigl(x^{[Q]}(u)-x^{[-Q]}(u)\bigr).
\]
Substituting the weak mirror branch of
Assumption~\ref{ass:konishi-leading-mirror-input} gives
\[
  \ii\widetilde p_Q(u)
  =
  -Q-2\ii g\left(
    \frac{g}{u+\ii Q/2}
    -\frac{u-\ii Q/2}{g}
  \right)+O(g^2)
  =
  2\ii u+O(g^2),
\]
and therefore
\begin{equation}
  \frac{\dd\widetilde p_Q}{\dd u}=2+O(g^2).
  \label{eq:konishi-leading-mirror-measure}
\end{equation}
This is the sign check that fixes the convention in the energy correction.

\begin{proposition}[Konishi weak-density rationalization]
\label{prop:konishi-weak-density-rationalization}
Let \(u_\ast=1/(2\sqrt3)\) and put \(q=2u\).  For \(Q\geq1\), define
\[
\begin{aligned}
  B_Q^-(q)
  &=
  9q^4+6(3(Q-2)Q+2)q^2+(3(Q-2)Q+4)^2,
\\
  B_Q^+(q)
  &=
  9q^4+6(3Q(Q+2)+2)q^2+(3Q(Q+2)+4)^2 .
\end{aligned}
\]
Then the weak mirror density \eqref{eq:konishi-yq0} is equivalently
\[
  Y_Q^{(0)}(q/2)
  =
  \frac{147456Q^2(3q^2+3Q^2-4)^2}
       {(q^2+Q^2)^4B_Q^-(q)B_Q^+(q)} .
  \label{eq:konishi-yq0-rationalized}
\]
\end{proposition}

\begin{proof}
The first denominator factor and the numerator become
\[
  u^2+\frac{Q^2}{4}=\frac{q^2+Q^2}{4},
  \qquad
  u^2-u_\ast^2+\frac{Q^2-1}{4}
  =
  \frac{3q^2+3Q^2-4}{12}.
\]
It remains only to rationalize the two factors involving
\(u_\ast\).  For any real \(a\),
\[
\begin{aligned}
  &\prod_{\varepsilon=\pm1}
  \left(\left(\frac q2-\frac{\varepsilon}{2\sqrt3}\right)^2+\frac{a^2}{4}\right)
\\
  &\qquad
  =
  \frac{1}{16}
  \left[
    q^4+2\left(a^2-\frac13\right)q^2
    +\left(a^2+\frac13\right)^2
  \right].
\end{aligned}
\]
Multiplying the bracket by \(9\), and setting \(a=Q-1\) and \(a=Q+1\),
gives respectively \(B_Q^-(q)\) and \(B_Q^+(q)\).  Hence each such product
is \(B_Q^\pm(q)/144\).  Substituting these three elementary identities in
\eqref{eq:konishi-yq0} gives
\[
  \frac{
    4Q^2\left((3q^2+3Q^2-4)/12\right)^2
  }{
    ((q^2+Q^2)/4)^4\,B_Q^-(q)B_Q^+(q)/144^2
  },
\]
which simplifies to \eqref{eq:konishi-yq0-rationalized}.
\end{proof}

\begin{lemma}[Pole structure of the Konishi weak density]
\label{lem:konishi-weak-density-poles}
Let \(Q\geq1\), and let \(Y_Q^{(0)}(q/2)\) be the rational function in
\eqref{eq:konishi-yq0-rationalized}.  Then
\[
\begin{aligned}
  B_Q^-(q)
  &=
  9\prod_{\sigma=\pm1}
  \left(\left(q-\frac{\sigma}{\sqrt3}\right)^2+(Q-1)^2\right),
\\
  B_Q^+(q)
  &=
  9\prod_{\sigma=\pm1}
  \left(\left(q-\frac{\sigma}{\sqrt3}\right)^2+(Q+1)^2\right).
\end{aligned}
\]
After the removable \(Q=1\) cancellation described below, the function has no
pole on the real \(q\)-axis and obeys \(Y_Q^{(0)}(q/2)=O(q^{-12})\) as
\(|q|\to\infty\).  Its upper-half-plane poles are the fourth-order pole
\(q=\ii Q\), the simple poles
\[
  q=\pm\frac1{\sqrt3}+\ii(Q+1),
\]
and, for \(Q>1\), the simple poles
\[
  q=\pm\frac1{\sqrt3}+\ii(Q-1).
\]
For \(Q=1\) the latter two points are real removable points, not contour
poles.
\end{lemma}

\begin{proof}
For any real \(a\),
\[
  \prod_{\sigma=\pm1}
  \left(\left(q-\frac{\sigma}{\sqrt3}\right)^2+a^2\right)
  =
  q^4+2\left(a^2-\frac13\right)q^2
  +\left(a^2+\frac13\right)^2 .
\]
Multiplication by \(9\), followed by \(a=Q-1\) and \(a=Q+1\), gives the two
displayed factorizations and therefore the zero locations.  The factor
\((q^2+Q^2)^4\) gives a fourth-order zero of the denominator at
\(q=\ii Q\) and at \(q=-\ii Q\).  For real \(q\), the factors with
\(a>0\) are strictly positive.  The only exceptional case is \(Q=1\) in
\(B_1^-\), where
\[
  B_1^-(q)=(3q^2-1)^2 .
\]
The numerator of \eqref{eq:konishi-yq0-rationalized} contains the identical
factor \((3q^2-1)^2\), so the singularities at \(q=\pm1/\sqrt3\) are
removable.  Finally, before this cancellation the numerator has degree
\(4\) and the denominator has degree \(16\).  For \(Q=1\), after cancellation
the numerator has degree \(0\) and the denominator has degree \(12\).  In both
cases the decay is \(q^{-12}\).
\end{proof}

\begin{lemma}[Residue reduction for the Konishi summand]
\label{lem:konishi-residue-reduction}
For \(Q\geq1\), put
\[
  \mathcal Y_Q(q)
  =
  \frac{147456Q^2(3q^2+3Q^2-4)^2}
       {(q^2+Q^2)^4B_Q^-(q)B_Q^+(q)}
\]
and
\[
  I_Q=-\int_{\mathbb R}\frac{\dd q}{2\pi}\,\mathcal Y_Q(q),
\]
with the \(Q=1\) removable real points understood as in
Lemma~\ref{lem:konishi-weak-density-poles}.  Then
\begin{equation}
  I_Q=R_Q+\frac{864}{Q^3}-\frac{1440}{Q^5},
  \label{eq:konishi-iq-summand}
\end{equation}
where
\[
  R_Q
  =
  -\frac{7776Q\,N(Q)}
  {(9Q^4-3Q^2+1)^4(27Q^6-27Q^4+36Q^2+16)}
\]
and
\[
\begin{aligned}
  N(Q)
  &=
  19683Q^{18}-78732Q^{16}+150903Q^{14}-134865Q^{12}
  +1458Q^{10}
\\
  &\quad
  +48357Q^8-13311Q^6-1053Q^4+369Q^2-10 .
\end{aligned}
\]
\end{lemma}

\begin{proof}
Write
\[
  \mathcal N_Q(q)=147456Q^2(3q^2+3Q^2-4)^2 .
\]
By Lemma~\ref{lem:konishi-weak-density-poles}, the upper contour encloses
the fourth-order pole at \(q=\ii Q\), the two simple poles with
\(a=Q+1\), and, when \(Q>1\), the two simple poles with \(a=Q-1\).  Hence
\[
  I_Q
  =
  -\ii\left(
    \rho_Q^{(4)}+\rho_Q^{(+)}+\mathbf 1_{Q>1}\rho_Q^{(-)}
  \right),
\]
where
\[
  \rho_Q^{(4)}
  =
  \frac1{3!}
  \left.
  \frac{\dd^3}{\dd q^3}
  \frac{\mathcal N_Q(q)}
       {(q+\ii Q)^4B_Q^-(q)B_Q^+(q)}
  \right|_{q=\ii Q}.
\]
For the simple poles, set \(a_\eta=Q+\eta\) for
\(\eta=\pm1\), \(q_{\sigma,\eta}=\sigma/\sqrt3+\ii a_\eta\), and use the
convention \(B_Q^{+1}=B_Q^+\), \(B_Q^{-1}=B_Q^-\).  When \(a_\eta>0\),
\[
  \rho_Q^{(\eta)}
  =
  \sum_{\sigma=\pm1}
  \frac{\mathcal N_Q(q_{\sigma,\eta})}
       {(q_{\sigma,\eta}^2+Q^2)^4
        B_Q^{-\eta}(q_{\sigma,\eta})
        \partial_q B_Q^\eta(q_{\sigma,\eta})}.
\]
The derivative in the denominator is not an implicit convention.  The
factorization in Lemma~\ref{lem:konishi-weak-density-poles} gives
\[
  \partial_qB_Q^\eta(q_{\sigma,\eta})
  =
  24\ii a_\eta(1+\sigma\ii\sqrt3\,a_\eta).
\]
These displayed formulas reduce the contour calculation to arithmetic in the
field \(\mathbb Q(\sqrt3,\ii)\).  Adding the \(\sigma=+1\) and \(\sigma=-1\)
residues cancels the \(\sqrt3\)-odd part.  Combining the two simple-pole
pairs with the fourth-order Laurent coefficient over a common denominator
then gives exactly \eqref{eq:konishi-iq-summand}.  The companion calculation
check performs this Laurent extraction literally, including the \(Q=1\)
omission of the removable \(a_{-}=0\) pair.
\end{proof}

\begin{proposition}[Leading Konishi wrapping integral]
\label{prop:konishi-leading-wrapping-integral}
Assume the mirror input of
Assumption~\ref{ass:konishi-leading-mirror-input}.  Then the order \(g^8\)
finite-size correction to the Konishi dimension is
\[
  \Delta_{\rm K}^{\rm wrap}
  =
  (324+864\zeta(3)-1440\zeta(5))g^8+O(g^{10}).
\]
\end{proposition}

\begin{proof}
Since \(Y_{\bullet_Q}=O(g^8)\),
\[
  \log(1+Y_{\bullet_Q})=Y_{\bullet_Q}+O(g^{16}).
\]
The \(O(g^{10})\) uncertainty in
\eqref{eq:konishi-leading-yq}, rather than the logarithm square, controls the
next uncomputed term.  Combining \eqref{eq:planar-n4-excited-tba-energy},
\eqref{eq:konishi-leading-yq}, and
\eqref{eq:konishi-leading-mirror-measure} gives
\[
  \Delta_{\rm K}^{\rm wrap}
  =
  -g^8\sum_{Q=1}^\infty
  \int_{\mathbb R}\frac{\dd u}{\pi}\,Y_Q^{(0)}(u)+O(g^{10}).
\]
Set \(q=2u\) and use
Proposition~\ref{prop:konishi-weak-density-rationalization}.  The coefficient
of \(g^8\) becomes
\[
  \sum_{Q=1}^\infty I_Q,
\]
with
\begin{equation}
  I_Q
  =
  -\int_{-\infty}^{\infty}\frac{\dd q}{2\pi}\,
  \frac{147456Q^2(3q^2+3Q^2-4)^2}
       {(q^2+Q^2)^4 B_Q^-(q)B_Q^+(q)},
  \label{eq:konishi-iq-integral}
\end{equation}
where \(B_Q^\pm\) are the polynomials in
Proposition~\ref{prop:konishi-weak-density-rationalization}.
By Lemma~\ref{lem:konishi-residue-reduction}, this integral satisfies
\eqref{eq:konishi-iq-summand}.
The non-zeta part is not left as a numerical experiment: it telescopes.  Define
\[
  F_Q
  =
  -\frac{648P(Q)}
  {(3Q^2+1)(3Q^2-6Q+4)(3Q^2-3Q+1)^4},
\]
with
\[
\begin{aligned}
  P(Q)
  &=
  486Q^{10}-2430Q^9+5184Q^8-6156Q^7+4752Q^6
  -2916Q^5
\\
  &\quad
  +1521Q^4-504Q^3+51Q^2+12Q-2 .
\end{aligned}
\]
A direct common-denominator calculation verifies
\[
  R_Q=F_Q-F_{Q+1}.
\]
Moreover \(F_1=324\) and \(F_Q\to0\) as \(Q\to\infty\).  Hence
\[
  \sum_{Q=1}^\infty R_Q=324.
\]
The remaining two sums in \eqref{eq:konishi-iq-summand} are absolutely
convergent and give \(864\zeta(3)-1440\zeta(5)\).  This proves the displayed
coefficient.
\end{proof}

Combining the asymptotic and wrapping pieces gives the physical four-loop
coefficient
\[
  -2496+576\zeta(3)-1440\zeta(5).
\]
The companion script
\path{calculation-checks/planar_n4_integrability_checks.py} now checks the
stringbook \(u\)-integrand against the \(q=2u\) rational form, numerically
integrates the first four real-line rational integrals, and verifies the
exact telescoping identity for the rational tail.  The conceptual assumptions
remain those isolated in
Assumption~\ref{ass:konishi-leading-mirror-input}; the finite residue and
summation algebra is not conjectural.

\section{Relation to Relativistic Integrable QFT}

The nested TBA of
Chapter~\ref{ch:integrable-nested-tba-tq-separation-variables} supplies the
finite-density variational logic.  The present system differs in three ways.
First, the dispersion is not relativistic.  Second, the mirror theory has an
infinite family of bound-state and auxiliary nodes.  Third, the spectral
problem is a gauge-theory local-operator problem, so the circle length is the
trace length \(L\) or a charge shifted from it by sector-dependent
conventions.  These differences are why the planar \(\mathcal N=4\) TBA is
placed in the supersymmetric gauge-theory volume rather than in the
relativistic integrable-QFT volume.
