\chapter{Three-Dimensional Chern--Simons--Matter Theories}
\label{ch:susy-three-dimensional-chern-simons-matter}

\ConventionsMixed

In three dimensions, Chern--Simons--matter theories combine topological gauge
terms, parity anomalies, monopole operators, and matter dynamics.  This chapter
treats them as field-theoretic data with specified definitions and duality
maps.  The first part fixes the local Chern--Simons and light-cone conventions,
including the vector-model large-\(N\) framework in which the gauge action
becomes quadratic.  The second part fixes the \(3D\) \(\mathcal N=2\)
Chern--Simons--matter conventions used by the localization chapter.  The final
part gives the Aharony--Bergman--Jafferis--Maldacena theory as a worked example
with gauge group, levels, matter, superpotential, conformal locus, moduli
space, monopole enhancement boundary, and \(S^3\) localization matrix model all
stated in the same normalization.

\section{Three-Dimensional Supersymmetry}

Let \(\eta_{\mu\nu}=\operatorname{diag}(-,+,+)\).  A minimal Majorana spinor
has two real components.  The \(\mathcal N=2\) algebra may be written with
complex supercharges \(Q_\alpha,\bar Q_\alpha\):
\[
  \{Q_\alpha,\bar Q_\beta\}
  =
  2(\gamma^\mu)_{\alpha\beta}P_\mu
  +2\ii\epsilon_{\alpha\beta}Z,
\]
where \(Z\) is a possible central charge.  Particle multiplets are
representations of this Hilbert-space algebra.  Superfields are off-shell
coordinate packages for local variables; the two notions are connected only
after dynamics and quantization have been specified.

A \(3D\) \(\mathcal N=2\) vector multiplet contains
\[
  (A_\mu,\sigma,\lambda,\bar\lambda,D),
\]
where \(A_\mu\) is a gauge field, \(\sigma\) is a real adjoint scalar,
\(\lambda,\bar\lambda\) are gauginos, and \(D\) is an auxiliary scalar.  A
chiral multiplet contains
\[
  (\phi,\psi,F)
\]
in a unitary representation of the gauge group.  These are field variables
in a chosen off-shell presentation.  A statement about BPS particles or local
operators must also specify the Hamiltonian, Gauss law, and operator
completion of the theory.

\section{Chern--Simons Datum}

Let \(G\) be compact and let \(P\to M_3\) be a principal \(G\)-bundle.  A
Chern--Simons action is
\[
  S_{\mathrm{CS}}[A]
  =
  \frac{k}{4\pi}
  \int_{M_3}\operatorname{Tr}
  \left(A\wedge dA+\frac23A\wedge A\wedge A\right).
\]
The trace convention and the allowed levels \(k\) are tied to the lattice of
large gauge transformations.  Gauge invariance of
\(\exp(\ii S_{\mathrm{CS}})\) requires the change of \(S_{\mathrm{CS}}\) under
a large gauge transformation to lie in \(2\pi\mathbb Z\).  For a spin
Chern--Simons theory the allowed counterterms may differ from the purely
bosonic theory by half-integral contact terms; this choice is part of the
three-dimensional QFT specification.

Varying the action gives
\[
  \delta S_{\mathrm{CS}}
  =
  \frac{k}{2\pi}\int_{M_3}\operatorname{Tr}(\delta A\wedge F_A)
  +
  \frac{k}{4\pi}\int_{\partial M_3}\operatorname{Tr}(A\wedge\delta A).
\]
On a closed manifold, the field equation of pure Chern--Simons theory is
\[
  F_A=0.
\]
On a manifold with boundary, a boundary condition or boundary degrees of
freedom are part of the gauge-theory specification.

\section{Conformal Chern--Simons--Matter Charts and Light-Cone Gauge}
\label{sec:cs-matter-light-cone-vector-models}

Chern--Simons--matter theories give a class of three-dimensional conformal
fixed points whose perturbative expansion is not organized by a dimensionful
Yang--Mills coupling.  The level \(k\) is quantized, while the perturbative
coordinate is a dimensionless parameter such as \(1/k\), or in vector models
the 't Hooft coordinate
\begin{equation}
  \lambda:=\frac{N}{k_{\rm eff}},
  \label{eq:cs-vector-thooft-coordinate}
\end{equation}
where \(k_{\rm eff}\) includes the parity-anomaly and regulator shift specified
with the theory.  This statement is not a claim that every Lagrangian written
with massless matter is automatically a conformal theory.  Relevant
deformations, such as scalar and fermion masses, must be tuned to the fixed
point, and scalar self-interactions must be placed at their fixed-point values
when they are present.  Once these data have been fixed, correlation functions
can be computed in ordinary diagrammatic perturbation theory around the
conformal point.

The light-cone gauge makes a special feature of Chern--Simons theory explicit.
Use
\begin{equation}
  x^\pm=\frac{x^0\pm x^1}{\sqrt2},\qquad x^\perp=x^2,\qquad
  ds^2=-2\,dx^+dx^-+(dx^\perp)^2,
  \label{eq:cs-light-cone-coordinates}
\end{equation}
and choose the orientation convention
\(\epsilon^{+-\perp}=+1\).  The gauge condition is
\begin{equation}
  A_-=0.
  \label{eq:cs-light-cone-gauge}
\end{equation}
The cubic Chern--Simons vertex then vanishes because every wedge product of
three one-forms in three dimensions contains one \(dx^-\).  Thus the pure gauge
part is first order and quadratic:
\begin{equation}
  S_{\rm CS}\big|_{A_-=0}
  =
  \frac{k}{2\pi}
  \int d^3x\,\operatorname{Tr}\!\left(A_+\partial_-A_\perp\right),
  \label{eq:cs-light-cone-quadratic-action}
\end{equation}
up to boundary terms and zero-mode data.

\paragraph{Component check of the light-cone Chern--Simons action.}
The component form of the local Chern--Simons density is
\[
  \frac{k}{4\pi}\epsilon^{\mu\nu\rho}
  \operatorname{Tr}\!\left(
    A_\mu\partial_\nu A_\rho
    +\frac23 A_\mu A_\nu A_\rho
  \right).
\]
The cubic term is antisymmetric in the three spacetime labels and therefore
contains one \(A_-\); it is zero after imposing \(A_-=0\).  The remaining
quadratic terms with no \(A_-\) are
\[
  \frac{k}{4\pi}
  \operatorname{Tr}\!\left(
    A_+\partial_-A_\perp
    -A_\perp\partial_-A_+
  \right),
\]
because \(\epsilon^{+-\perp}=+1\) and
\(\epsilon^{\perp-+}=-1\).  Integrating the second term by parts in \(x^-\)
and dropping the boundary term gives twice
\(\operatorname{Tr}(A_+\partial_-A_\perp)\), giving
\eqref{eq:cs-light-cone-quadratic-action}.

Let matter in a representation \(R\) of \(G\) define the gauge current
\[
  J^{\mu a}:=-\frac{\delta S_{\rm matter}}{\delta A_\mu^a}
\]
in the same connection normalization as above.  At finite light-front
regulator, remove the \(\partial_-\) zero mode or keep it as a separate
Gauss-law degree of freedom.  On the zero-mode-free subspace write
\(\partial_-^{-1}\) for the inverse with the declared prescription.

\paragraph{Light-cone Chern--Simons current kernel.}
At finite regulator, the light-cone-gauge integration over \(A_+\) and
\(A_\perp\) is an ordinary Gaussian integration on the nonzero-mode subspace.
With the source convention
\begin{equation}
  S[A_+,A_\perp;J]
  =
  \int d^3x\left[
  \frac{k}{2\pi}A_+^a\partial_-A_\perp^a
  -A_+^aJ^{+a}-A_\perp^aJ^{\perp a}\right],
  \label{eq:cs-light-cone-source-action}
\end{equation}
substitution of the classical solution gives the effective current interaction
\begin{equation}
  S_{\rm eff}[J]
  =
  -\frac{2\pi}{k}
  \int d^3x\,
  J^{+a}\,\partial_-^{-1}J^{\perp a},
  \label{eq:cs-light-cone-effective-current}
\end{equation}
with the ordering and zero-mode prescription displayed.

Set \(a=k/(2\pi)\) and \(D=\partial_-\).  The equations following from
\eqref{eq:cs-light-cone-source-action} are
\[
  aD A_\perp^a=J^{+a},\qquad
  aD A_+^a=-J^{\perp a},
\]
where the second equation uses integration by parts in \(x^-\).  Hence
\[
  A_\perp^a=a^{-1}D^{-1}J^{+a},
  \qquad
  A_+^a=-a^{-1}D^{-1}J^{\perp a}.
\]
Substituting into the action, the bilinear term cancels the \(A_+J^+\) source
term and leaves
\[
  -A_\perp^aJ^{\perp a}
  =
  -a^{-1}(D^{-1}J^{+a})J^{\perp a}.
\]
This is \eqref{eq:cs-light-cone-effective-current}.  The calculation is a
finite-dimensional completion of the square after the regulator has replaced
the nonzero-mode function space by a finite or Hilbert-space Gaussian problem.
The zero mode is not determined by this inverse; it is the global Gauss-law
constraint and must be supplied as part of the light-front regulator.

This is the three-dimensional analogue of the simplification used in the
two-dimensional large-\(N\) 't Hooft model of
Chapter~\ref{ch:large-n-two-dimensional-qcd-light-front}.  The kernel is
\(\partial_-^{-1}\), not \((-\partial_-^2)^{-1}\), because the Chern--Simons
kinetic operator is first order rather than Yang--Mills second order.  In both
cases the light-front gauge field has no transverse propagating polarization
and imposes a nonlocal current interaction after the constraint is solved.

\begin{definition}[Planar Chern--Simons vector-model datum]
\label{def:planar-cs-vector-model-datum}
A planar Chern--Simons vector-model datum consists of:
\begin{enumerate}[label=(\roman*)]
\item gauge group \(U(N)\) or \(SU(N)\), global form, spin structure, framing
  convention, and effective level \(k_{\rm eff}\);
\item fundamental matter fields, such as scalars or Dirac fermions, with all
  relevant deformations tuned to the conformal point;
\item a parity-anomaly regulator fixing \(k_{\rm eff}\) and all background
  Chern--Simons contact terms;
\item a large-\(N\) limit with \(\lambda=N/k_{\rm eff}\) fixed;
\item a light-front or Euclidean regulator specifying zero modes,
  \(\partial_-^{-1}\), and the renormalized gauge-invariant singlet
  operators whose correlators are computed.
\end{enumerate}
\end{definition}

For fundamental matter the color algebra uses the same trace-delta
normalization as the QCD chapters,
\begin{equation}
  (t^a)^i{}_j(t^a)^k{}_\ell
  =
  \delta^i{}_\ell\delta^k{}_j
  -
  \frac1N\delta^i{}_j\delta^k{}_\ell .
  \label{eq:cs-vector-fundamental-completeness}
\end{equation}
Consequently the interaction
\eqref{eq:cs-light-cone-effective-current} preserves the single color trace at
leading order, while the second term in
\eqref{eq:cs-vector-fundamental-completeness} is suppressed by \(1/N\) in
connected singlet correlators.  The trace-preserving leading contraction is the
planar color algebra behind the bilocal closure of the two-dimensional
't Hooft model and the singlet light-front closure used here.

\subsection{Finite-Regulator Singlet Closure}

The preceding statement can be written as a finite-dimensional algebraic
identity before taking the large-\(N\) limit.  Let \(\Psi_i^A\) denote the
regulated fundamental matter variables that remain after gauge fixing; the
index \(i=1,\ldots,N\) is color, while \(A\) collects spin, flavor, position
or momentum, and any regulator labels.  Let \(\overline\Psi_A^i\) be the
conjugate variable in the antifundamental representation.  Define the singlet
bilocal matrix
\begin{equation}
  B^A{}_{B}
  =
  \frac1N\overline\Psi^i_B\Psi_i^A .
  \label{eq:cs-vector-bilocal-singlet}
\end{equation}
A gauge current component that is bilinear in the matter variables has the
form
\[
  J^{\mu a}
  =
  \overline\Psi^i_A\,(V^\mu)^A{}_{B}\,(t^a)^j{}_{i}\,\Psi_j^B ,
\]
where \(V^\mu\) is the matrix, including all differential or regulator
kernels, carried by the current component.  The displayed algebraic identity
below is written for ordinary complex variables.  In fermionic vector models
the current product is instead a graded product in the declared current order;
the color contraction and its \(N\)-scaling are identical, while signs coming
from moving Grassmann variables belong to the spinor and kernel convention of
the particular Schwinger--Dyson equation rather than to the color algebra.

\paragraph{Planar color reduction of the light-cone kernel.}
Let \(K\) be any regulator kernel acting on the non-color labels; in the
light-cone Chern--Simons problem \(K=\partial_-^{-1}\) with the zero-mode
prescription already declared.  With the normalization
\eqref{eq:cs-vector-fundamental-completeness},
\begin{align}
  J^{+a}KJ^{\perp a}
  &=
  N^2\,
  (V^+)^A{}_{B}\,B^B{}_{C}\,
  K^C{}_{D}\,
  (V^\perp)^D{}_{E}\,B^E{}_{A}
  \nonumber\\
  &\quad
  -N\,
  (V^+)^A{}_{B}\,B^B{}_{A}\,
  K^C{}_{D}\,
  (V^\perp)^D{}_{E}\,B^E{}_{C}.
  \label{eq:cs-vector-bilocal-color-reduction}
\end{align}
Consequently the light-cone interaction
\eqref{eq:cs-light-cone-effective-current} has the form
\[
  S_{\rm eff}
  =
  -2\pi N\lambda\,
  (V^+)^A{}_{B}B^B{}_{C}K^C{}_{D}(V^\perp)^D{}_{E}B^E{}_{A}
  +
  O(N^0)
\]
at fixed \(\lambda=N/k_{\rm eff}\).

Insert the current expression and use
\eqref{eq:cs-vector-fundamental-completeness}:
\[
  (t^a)^j{}_{i}(t^a)^\ell{}_{k}
  =
  \delta^j{}_{k}\delta^\ell{}_{i}
  -
  \frac1N\delta^j{}_{i}\delta^\ell{}_{k}.
\]
The first color contraction produces
\[
  \overline\Psi^i_A(V^+)^A{}_{B}\Psi_k^B\,
  K^C{}_{D}\,
  \overline\Psi^k_C(V^\perp)^D{}_{E}\Psi_i^E
  =
  N^2
  (V^+)^A{}_{B}B^B{}_{C}K^C{}_{D}(V^\perp)^D{}_{E}B^E{}_{A}.
\]
The trace-subtraction term gives the second line of
\eqref{eq:cs-vector-bilocal-color-reduction} with one fewer factor of \(N\).
Multiplication by the kernel coefficient \(2\pi/k_{\rm eff}=2\pi\lambda/N\)
then makes the first term order \(N\) and the trace-subtraction term order
\(N^0\).

\begin{proposition}[Finite-regulator Schwinger--Dyson identity]
\label{prop:cs-vector-finite-regulator-sd}
At finite regulator, let \(S_{\rm reg}[\Psi,\overline\Psi]\) be the matter
action after the light-cone gauge field has been integrated out as in
\eqref{eq:cs-light-cone-effective-current}.  For an ordinary complex bosonic
fundamental vector, integration by parts gives
\begin{equation}
  \left\langle
  \frac{\partial \mathcal O}{\partial \Psi_i^A}
  -
  \mathcal O\,\frac{\partial S_{\rm reg}}{\partial \Psi_i^A}
  \right\rangle_{\rm reg}
  =
  0
  \label{eq:cs-vector-bosonic-sd}
\end{equation}
for every polynomial singlet observable \(\mathcal O\).  For a Grassmann
fundamental vector the same equation holds with the ordinary derivative
replaced by the left Berezin derivative and with the standard graded sign
when the derivative crosses \(\mathcal O\).  If the large-\(N\) singlet sector
has factorization
\begin{equation}
  \langle B_1\cdots B_r\rangle_{\rm conn}=O(N^{1-r}),
  \label{eq:cs-vector-factorization}
\end{equation}
then the leading equations obtained from
\eqref{eq:cs-vector-bosonic-sd} close on the one-bilocal expectation
\(G^A{}_{B}:=\langle B^A{}_{B}\rangle\).
\end{proposition}

\begin{proof}
At finite regulator the bosonic statement is the divergence theorem applied to
the finite-dimensional integral over the variables \(\Psi_i^A\), with boundary
terms removed by the regulator damping or by compact contour deformation:
\[
  0
  =
  \int d\Psi\,d\overline\Psi\,
  \frac{\partial}{\partial\Psi_i^A}
  \left(\mathcal O\,\ee^{-S_{\rm reg}}\right).
\]
Expanding the derivative gives
\eqref{eq:cs-vector-bosonic-sd}.  The Grassmann version is the corresponding
translation identity for the Berezin integral.  After the gauge field has
been integrated out, the bilocal color reduction above shows that the
order-\(N\) interaction is a functional only of \(B\).  Taking
\(\mathcal O\) to be a product of singlet bilocals therefore gives equations
whose leading terms contain products of \(B\)'s.  The factorization estimate
\eqref{eq:cs-vector-factorization} replaces those products, at leading order,
by products of \(G\).  Thus no nonsinglet correlator enters the leading
large-\(N\) Schwinger--Dyson equation.
\end{proof}

\subsection{Planar Bilocal Saddle and Self-Energy Equation}

The preceding closure becomes computational only after the leading planar
equation is written as an equation for the exact two-point function.  We give
the finite-regulator derivation because it is the step often hidden behind
the phrase ``rainbow resummation'' in the Chern--Simons vector-model
literature.

Let \(Q^A{}_{B}\) be the regulated inverse free propagator of the fundamental
matter variable in the gauge-fixed theory, including all spinor, flavor,
momentum, and regulator labels but no color index.  Thus the quadratic part
of the action is
\[
  S_0=\overline\Psi^i_A Q^A{}_{B}\Psi_i^B .
\]
Keeping the \(O(N)\) term of
the planar color reduction above and placing the
\(O(N^0)\) trace-subtraction in the \(1/N\) expansion, the leading
light-cone interaction is
\begin{equation}
  S_{\rm int}^{(0)}
  =
  N\,I_{\rm LC}[B],
  \qquad
  I_{\rm LC}[B]
  =
  -2\pi\lambda\,
  (V^+)^A{}_{B}B^B{}_{C}
  K^C{}_{D}
  (V^\perp)^D{}_{E}B^E{}_{A},
  \label{eq:cs-vector-leading-bilocal-interaction}
\end{equation}
where \(K=\partial_-^{-1}\) with the same zero-mode prescription as in
\eqref{eq:cs-light-cone-effective-current}.  Define the color-normalized
two-point bilocal
\[
  G^A{}_{B}
  =
  \left\langle B^A{}_{B}\right\rangle
  =
  \frac1N\left\langle\overline\Psi^i_B\Psi_i^A\right\rangle .
\]

\begin{proposition}[Planar self-energy equation from the bilocal action]
\label{prop:cs-vector-planar-self-energy-equation}
Assume the finite-regulator planar vector-model datum of
Definition~\ref{def:planar-cs-vector-model-datum}, the leading interaction
\eqref{eq:cs-vector-leading-bilocal-interaction}, and large-\(N\) singlet
factorization in the form
\eqref{eq:cs-vector-factorization}.  For ordinary complex bosonic fundamental
variables, the leading large-\(N\) two-point function obeys
\begin{equation}
  G^A{}_{C}\,
  \bigl(Q^C{}_{B}+\Sigma^C{}_{B}[G]\bigr)
  =
  \delta^A{}_{B},
  \label{eq:cs-vector-planar-dyson-equation}
\end{equation}
where the self-energy is the functional derivative of the leading bilocal
interaction:
\begin{equation}
  \Sigma^C{}_{M}[G]
  =
  \frac{\partial I_{\rm LC}}{\partial G^M{}_{C}}
  =
  -2\pi\lambda
  \left[
    (V^+)^A{}_{M}K^C{}_{D}(V^\perp)^D{}_{E}G^E{}_{A}
    +
    (V^+)^C{}_{B}G^B{}_{D}K^D{}_{E}(V^\perp)^E{}_{M}
  \right].
  \label{eq:cs-vector-planar-self-energy-functional}
\end{equation}
For fermionic vector models the same color and bilocal derivative formula is
read in the declared graded order; the additional signs are those of the
Berezin Schwinger--Dyson identity and of the spinor current matrices
\(V^\mu\).
\end{proposition}

\begin{proof}
Apply the finite-regulator Schwinger--Dyson identity
\eqref{eq:cs-vector-bosonic-sd} with
\(\mathcal O=\Psi_i^A\).  Summing over the color label and dividing by \(N\)
gives
\[
  \delta^A{}_{M}
  =
  \frac1N
  \left\langle
    \Psi_i^A
    \frac{\partial S_{\rm reg}}{\partial\Psi_i^M}
  \right\rangle .
\]
The quadratic action contributes
\[
  \frac1N
  \left\langle\Psi_i^A\overline\Psi^i_C\right\rangle Q^C{}_{M}
  =
  G^A{}_{C}Q^C{}_{M}.
\]
Here the displayed equality uses that the variables in the proposition are
ordinary bosonic complex variables, so the two fields in the finite
regulator integral may be interchanged without a Koszul sign.
For the leading planar interaction, the chain rule and the definition of
\(B\) give
\[
  \frac{\partial}{\partial\Psi_i^M}
  \bigl(NI_{\rm LC}[B]\bigr)
  =
  \overline\Psi^i_C
  \frac{\partial I_{\rm LC}}{\partial B^M{}_{C}} .
\]
Therefore the interaction contribution is
\[
  \frac1N
  \left\langle
    \Psi_i^A\overline\Psi^i_C
    \frac{\partial I_{\rm LC}}{\partial B^M{}_{C}}
  \right\rangle .
\]
Large-\(N\) factorization replaces the product of the bilocal and the
bilocal polynomial
\(\partial I_{\rm LC}/\partial B^M{}_{C}\) by the same expression evaluated
at \(B=G\), up to \(O(N^{-1})\) corrections in this normalization.  This gives
\[
  \delta^A{}_{M}
  =
  G^A{}_{C}
  \left(
    Q^C{}_{M}
    +
    \left.
    \frac{\partial I_{\rm LC}}{\partial B^M{}_{C}}
    \right|_{B=G}
  \right),
\]
which is \eqref{eq:cs-vector-planar-dyson-equation}.

It remains only to compute the derivative of
\eqref{eq:cs-vector-leading-bilocal-interaction}.  Differentiating the first
and second occurrence of \(B\) gives respectively
\[
  -2\pi\lambda
  (V^+)^A{}_{M}K^C{}_{D}(V^\perp)^D{}_{E}G^E{}_{A},
  \qquad
  -2\pi\lambda
  (V^+)^C{}_{B}G^B{}_{D}K^D{}_{E}(V^\perp)^E{}_{M}.
\]
Their sum is \eqref{eq:cs-vector-planar-self-energy-functional}.  The
fermionic statement follows from the finite-regulator Berezin
Schwinger--Dyson identity with the same color contraction; one must not infer
the graded signs from the bosonic display without fixing the fermion-current
ordering.
\end{proof}

In a translation-invariant continuum chart, the labels \(A,B\) include
momenta and spin.  Equation~\eqref{eq:cs-vector-planar-dyson-equation} then
becomes the familiar planar Schwinger--Dyson integral equation for the exact
fundamental propagator.  Solving it requires the regulator-dependent mass and
contact counterterms that tune the theory to the conformal point.  The
mathematical point of the derivation above is that the planar self-energy is
the leading finite-regulator Schwinger--Dyson equation for gauge-invariant
singlet bilocals after the light-cone Chern--Simons constraint has been
integrated out.

\begin{controlledapproximation}[Planar light-front closure]
\label{ca:planar-cs-vector-light-front-closure}
In the datum of Definition~\ref{def:planar-cs-vector-model-datum}, the
large-\(N\) light-front perturbation series for single-trace singlet
correlators is controlled by planar ladders and rainbow self-energies built
from the kernel \eqref{eq:cs-light-cone-effective-current}.  The closure is a
statement about the declared regulator and the singlet sector: it gives
integral equations for exact planar two- and three-point functions once the
mass and contact terms have been tuned to the conformal point.  It is not a
proof of nonperturbative existence of the continuum theory and does not identify
colored fundamental fields as asymptotic states.
\end{controlledapproximation}

The useful lesson for the monograph is conceptual as well as computational.
Chern--Simons vector models are conformal gauge theories in which conventional
diagrams can be evaluated at the fixed point, and light-cone gauge converts the
topological gauge field into an exactly solvable constraint kernel.  This makes
them a higher-dimensional counterpart of the 't Hooft model: not identical in
dynamics, because the kernel, spin, parity anomaly, and operator spectrum are
different, but similar in the precise sense that the gauge field can be
integrated out before solving the planar singlet equations.

\section{Supersymmetric Chern--Simons--Matter Action}

The supersymmetric Chern--Simons action contains
\begin{equation}
  S_{\mathrm{CS}}^{\mathcal N=2}
  =
  \frac{k}{4\pi}\int
  \operatorname{Tr}
  \left(
    A\wedge dA+\frac23A^3
    -\bar\lambda\lambda
    +2D\sigma
  \right),
  \label{eq:three-d-susy-cs-vector-action}
\end{equation}
with the displayed signs defining the convention of this chapter.  Let
chiral multiplets \(\Phi_i\) take values in a unitary representation \(R\) of
\(G\).  Their scalar fields are \(\phi_i\).  With real mass parameters
\(m_i\), FI coordinate \(\zeta\) for an abelian factor, and superpotential
\(W\), the bosonic potential contains
\begin{equation}
  \sum_i\left|(\sigma+m_i)\phi_i\right|^2
  +
  \frac12\left(\mu(\phi)+\frac{k}{2\pi}\sigma-\zeta\right)^2
  +
  |dW(\phi)|^2,
  \label{eq:three-d-n2-bosonic-potential}
\end{equation}
in an abelian normalization where \(\mu\) is the matter moment map.
Supersymmetric classical vacua solve
\begin{equation}
  dW=0,\qquad
  (\sigma+m_i)\phi_i=0,\qquad
  \mu(\phi)+\frac{k}{2\pi}\sigma-\zeta=0,
  \label{eq:three-d-n2-vacuum-equations}
\end{equation}
modulo gauge transformations.

\begin{remark}[Auxiliary-field origin of the shifted D-term]
\label{rem:three-d-cs-d-term-shift}
In the abelian convention of \eqref{eq:three-d-susy-cs-vector-action}, the
auxiliary equation shifts the moment-map equation by
\[
  \mu(\phi)-\zeta\longmapsto
  \mu(\phi)+\frac{k}{2\pi}\sigma-\zeta .
\]

The part of the Lagrangian involving \(D\) has the form
\[
  \frac12D^2+D\bigl(\mu(\phi)-\zeta\bigr)
  +\frac{k}{2\pi}D\sigma ,
\]
where the last term is the \(2D\sigma\) term in
\eqref{eq:three-d-susy-cs-vector-action}.  Completing the square gives
\[
  \frac12
  \left(D+\mu(\phi)+\frac{k}{2\pi}\sigma-\zeta\right)^2
  -
  \frac12
  \left(\mu(\phi)+\frac{k}{2\pi}\sigma-\zeta\right)^2 .
\]
Eliminating \(D\) leaves the displayed positive potential and hence the
vacuum equation \eqref{eq:three-d-n2-vacuum-equations}.
\end{remark}

\paragraph{Non-abelian Chern--Simons auxiliary elimination.}
Let \(G\) be compact, let \(R\) be a unitary representation, and choose
Hermitian representation matrices \(t_R^a\) with trace convention
\(\operatorname{Tr}(t^at^b)=\delta^{ab}\) for the gauge action.  Assume
\(k\neq 0\), no Yang--Mills kinetic term, no FI term, no real mass, and no
superpotential.  For one chiral multiplet in \(R\), define the finite
moment-map coordinates
\[
  \mu^a(\phi)=\overline\phi\,t_R^a\phi .
\]
For several chiral multiplets, replace \(R\) by the direct sum of their
representations, so that \(\mu^a\) is the sum of the corresponding quadratic
coordinates.
Then the local non-derivative bosonic component terms involving the
Chern--Simons auxiliary scalar \(\sigma=\sigma^a t^a\) and the auxiliary
field \(D=D^a t^a\) may be written as
\begin{equation}
  \mathcal L_{\mathrm{bos,aux}}
  =
  \frac{k}{2\pi}D^a\sigma^a
  +
  D^a\mu^a(\phi)
  -
  \sigma^a\sigma^b\,
  \overline\phi\, t_R^a t_R^b\phi .
  \label{eq:three-d-n2-nonabelian-auxiliary-lagrangian}
\end{equation}
Eliminating \(D\) imposes
\begin{equation}
  \sigma^a=-\frac{2\pi}{k}\mu^a(\phi),
  \label{eq:three-d-n2-cs-sigma-solution}
\end{equation}
and the remaining scalar interaction is
\begin{equation}
  \mathcal L_{\mathrm{bos,on\text{-}shell}}
  =
  -\frac{4\pi^2}{k^2}
  \bigl(\overline\phi\,t_R^a\phi\bigr)
  \bigl(\overline\phi\,t_R^b\phi\bigr)
  \bigl(\overline\phi\,t_R^a t_R^b\phi\bigr).
  \label{eq:three-d-n2-cs-sextic-scalar}
\end{equation}
Equivalently, the scalar potential contains the positive contribution
\[
  V_{\mathrm{sextic}}
  =
  \frac{4\pi^2}{k^2}
  \bigl(\overline\phi\,t_R^a\phi\bigr)
  \bigl(\overline\phi\,t_R^b\phi\bigr)
  \bigl(\overline\phi\,t_R^a t_R^b\phi\bigr)
\]
in the same local coordinates.

The field \(D\) is a Lagrange multiplier in the pure Chern--Simons
normalization.  Varying
\eqref{eq:three-d-n2-nonabelian-auxiliary-lagrangian} with respect to
\(D^a\) gives
\[
  \frac{k}{2\pi}\sigma^a+\mu^a(\phi)=0,
\]
which is \eqref{eq:three-d-n2-cs-sigma-solution}.  Substitution into the last
term of \eqref{eq:three-d-n2-nonabelian-auxiliary-lagrangian} gives
\[
  -\left(-\frac{2\pi}{k}\mu^a\right)
   \left(-\frac{2\pi}{k}\mu^b\right)
   \overline\phi\,t_R^at_R^b\phi
  =
  -\frac{4\pi^2}{k^2}
   \mu^a\mu^b\overline\phi\,t_R^at_R^b\phi .
\]
No commutativity of the matrices \(t_R^a\) and \(t_R^b\) has been used; the
ordered product in \eqref{eq:three-d-n2-cs-sextic-scalar} is part of the
component convention.

The same finite-dimensional elimination fixes the power of \(k\) and the
relative sizes of the induced Yukawa terms.  The signs of the terms carrying
an explicit \(\ii\) depend on the three-dimensional spinor conjugation and on
where the \(\ii\) is placed in the Chern--Simons gaugino mass.  In the common
component convention in which the non-derivative fermion terms are normalized
as in the stringbook appendices, one obtains coefficient magnitudes
\[
  \frac{2\pi}{\lvert k\rvert}
  \quad\hbox{for}\quad
  \bigl(\overline\phi\,t_R^a\phi\bigr)
  \bigl(\overline\psi\,t_R^a\psi\bigr),
  \qquad
  \frac{4\pi}{\lvert k\rvert}
  \quad\hbox{for}\quad
  \bigl(\overline\psi\,t_R^a\phi\bigr)
  \bigl(\overline\phi\,t_R^a\psi\bigr).
\]
Thus the convention-independent check is the \(1:2\) relative coefficient and
the single inverse power of \(k\); the signed formula must be read only after
the spinor bilinear convention has been fixed.

\begin{remark}[\(\mathcal N=3\) adjoint-chiral elimination]
\label{rem:three-d-n3-adjoint-chiral-elimination}
Let \(Q\) and \(\widetilde Q\) be conjugate chiral multiplets in
representations \(R\) and \(R^\vee\), and let \(\varphi=\varphi^a t^a\) be
an adjoint chiral multiplet.  With the same trace convention as above, set
\[
  J^a=\widetilde Q\,t_R^a Q .
\]
The \(\mathcal N=3\) Chern--Simons-matter superpotential
\[
  W(\varphi,Q,\widetilde Q)
  =
  -\frac{k}{8\pi}\varphi^a\varphi^a
  +
  J^a\varphi^a
\]
has algebraic \(\varphi\)-equation
\[
  \varphi^a=\frac{4\pi}{k}J^a,
\]
and the exact tree-level effective superpotential after eliminating
\(\varphi\) is
\[
  W_{\mathrm{eff}}(Q,\widetilde Q)
  =
  \frac{2\pi}{k}J^aJ^a
  =
  \frac{2\pi}{k}
  \bigl(\widetilde Q\,t_R^aQ\bigr)
  \bigl(\widetilde Q\,t_R^aQ\bigr).
\]

The F-term equation is
\[
  -\frac{k}{4\pi}\varphi^a+J^a=0.
\]
Therefore \(\varphi^a=4\pi J^a/k\).  Substitution gives
\[
  -\frac{k}{8\pi}
  \left(\frac{4\pi}{k}\right)^2J^aJ^a
  +
  \frac{4\pi}{k}J^aJ^a
  =
  -\frac{2\pi}{k}J^aJ^a
  +
  \frac{4\pi}{k}J^aJ^a
  =
  \frac{2\pi}{k}J^aJ^a .
\]
The calculation is an elimination in the polynomial chiral ring, so it does
not use a string-theoretic realization of the same Lagrangian.
\end{remark}

\section{Parity Anomaly And Contact Terms}

A massive three-dimensional Dirac fermion in representation \(R\) shifts the
effective Chern--Simons level by
\[
  \Delta k=\frac12\,\operatorname{sign}(m)\,T(R),
\]
where
\[
  \operatorname{Tr}_R(t^at^b)=T(R)\delta^{ab}
\]
in the trace convention used for the gauge action.  Therefore a massless
fermion coupled to a gauge field requires a choice of regulator that fixes
the integer or half-integer effective level.  This choice is part of the
definition of the continuum theory and of its time-reversal or parity
properties.

Background flavor and R-symmetry Chern--Simons terms are equally important.
Two proposed dual theories with the same local operator spectrum can still
differ as QFTs if their background contact terms differ.  In localization
formulae these contact terms appear as phases or Gaussian factors in the
Cartan variables; they are not disposable normalizations.

\section{Monopole Operators}

Let \(G\) have maximal torus \(T\).  A monopole operator at the origin is
defined by imposing a magnetic singularity on a small sphere:
\[
  \frac1{2\pi}\int_{S^2}F=m\in\operatorname{Hom}(U(1),T)/W_G .
\]
The local operator is gauge invariant only after its electric charges induced
by Chern--Simons terms and fermion zero modes are accounted for.  In an
abelian theory with level \(k\), Gauss law gives electric charge
\[
  q_{\mathrm{el}}=k\,m
\]
for a bare monopole of magnetic charge \(m\).  Gauge-invariant monopole
operators may require dressing by charged matter fields.

This point is not optional in supersymmetric Chern--Simons--matter theory.
Monopole operators can carry flavor and R-charges, can complete multiplets
that are invisible in a polynomial superfield presentation, and can change
the apparent global symmetry of the interacting fixed point.  Thus a
Lagrangian written in \(\mathcal N=2\) variables may describe a theory with
larger supersymmetry only after the monopole sectors are included.

\section{ABJM Gauge-Theory Specification}
\label{sec:abjm-gauge-theory-specification}

Fix integers \(N\ge1\) and \(k\ne0\).  The ABJM gauge group and levels are
\[
  G_{N,k}=U(N)_k\times U(N)_{-k},
\]
with connections \(A\) and \(\widehat A\).  The subscript means that the
Chern--Simons functional of \(A\) has level \(k\), while that of
\(\widehat A\) has level \(-k\), in the same trace normalization.  The
global form of the gauge group and the quotient by any diagonal one-form
symmetry affect line operators and must be declared in a fully specified
theory.  In this chapter \(U(N)\times U(N)\) denotes the standard ABJM
choice used for the local-operator and \(S^3\) partition-function formulae.

In \(\mathcal N=2\) language the matter consists of four chiral multiplets
\[
  A_a\in(\mathbf N,\overline{\mathbf N}),\qquad
  B_{\dot a}\in(\overline{\mathbf N},\mathbf N),
  \qquad a,\dot a=1,2 .
\]
The classical superpotential is
\begin{equation}
  W_{\rm ABJM}
  =
  \frac{2\pi}{k}
  \varepsilon^{ab}\varepsilon^{\dot a\dot b}
  \operatorname{Tr}
  \bigl(A_aB_{\dot a}A_bB_{\dot b}\bigr).
  \label{eq:abjm-superpotential}
\end{equation}
The trace is over the first \(U(N)\) gauge indices, so every product in
\eqref{eq:abjm-superpotential} is an endomorphism of the first gauge factor.
The displayed coefficient fixes the convention in which the Yukawa and
sextic scalar terms assemble into the \(\mathcal N=6\) invariant potential
after auxiliary fields are eliminated.  The \(\mathcal N=2\) trial
\(R\)-charge assignment
\[
  R(A_a)=R(B_{\dot a})=\frac12
\]
gives \(R(W_{\rm ABJM})=2\), as required for a superpotential term.

The scalar variables can be grouped as
\[
  C_I=(A_1,A_2,B_1^\dagger,B_2^\dagger),
  \qquad I=1,\ldots,4 .
\]
The interacting theory has an \(SU(4)\) flavor symmetry rotating the \(C_I\)
in the \(\mathcal N=6\) presentation.  This is a field-theory statement
about the completed gauge theory, including monopole sectors and the
Chern--Simons Gauss law.  The holomorphic \(\mathcal N=2\) superpotential
variables display only one coordinate presentation of this symmetry.

\paragraph{Parity and the opposite levels.}
Let \(P\) be an orientation-reversing diffeomorphism of the three-manifold.
The Chern--Simons functional changes sign under \(P\).  Therefore the pair of
levels \((k,-k)\) is invariant under the combined operation
\[
  P:\quad A\longleftrightarrow\widehat A,\qquad
  A_a\longleftrightarrow B_{\dot a}^{\dagger},
\]
up to the corresponding exchange of flavor indices and spinor conventions:
under an orientation reversal the integral of a three-form changes sign, so
\[
  S_{\rm CS,k}[A]\mapsto -S_{\rm CS,k}[P^\ast A].
\]
Exchanging the two gauge factors sends the level \(k\) contribution to the
sector that originally carried level \(-k\), and conversely.  Hence the sum
\[
  S_{\rm CS,k}[A]+S_{\rm CS,-k}[\widehat A]
\]
is mapped to itself after relabeling the two connections.  The matter
exchange identifies the bifundamental and antibifundamental representations,
so the gauge-covariant kinetic terms and the quartic superpotential are
carried into their complex conjugate form with the same real action.

\section{ABJM Conformal Locus And Marginal Couplings}
\label{sec:abjm-conformal-locus}

The ABJM family is often described as if a Chern--Simons level were an
ordinary coupling.  For conformal-manifold questions this is the wrong
category: the level is part of the topological gauge specification and is quantized
by large gauge transformations.  We first state the local notion being used.

\begin{definition}[Local conformal manifold in a fixed QFT specification]
\label{def:fixed-qft-specification-conformal-manifold}
Fix a \(d\)-dimensional continuum QFT specification \(\mathfrak D\), including the
global form of the gauge group, line-operator lattice, spin and framing
choices, background contact-term scheme, and the global or supersymmetry
algebra required to be preserved.  A local conformal manifold through a
conformal theory \(\mathcal T_0\) with specification \(\mathfrak D\) is a germ of a
smooth parameter space \(U\) and a family \(\{\mathcal T_u\}_{u\in U}\) of
conformal QFTs with the same specification such that tangent vectors at \(0\) are
represented by integrated scalar primary deformations
\[
  \delta S
  =
  \sum_a \delta u^a\int d^d x\,\mathcal O_a(x),
  \qquad
  \Delta(\mathcal O_a)=d,
\]
modulo redundant deformations coming from total derivatives, symmetry
variations, and field-coordinate changes.  A marginal Lagrangian parameter
is exactly marginal only after it defines such a tangent vector with all beta
functions and background contact terms remaining inside \(\mathfrak D\).
\end{definition}

\begin{hypothesis}[Standard ABJM conformal class]
\label{hyp:standard-abjm-conformal-class}
In this section ``standard ABJM'' means the local theory with gauge specification
\[
  U(N)_k\times U(N)_{-k},\qquad k\in\mathbb Z\setminus\{0\},
\]
the four bifundamental chiral multiplets of
Section~\ref{sec:abjm-gauge-theory-specification}, zero real masses, zero FI
coordinates, the standard line-operator lattice used in the \(S^3\)
localization formula, and the interacting \(\mathcal N=6\) superconformal
symmetry with \(SU(4)\) flavor symmetry.  Within this class the quartic
superpotential tensor is the one in \eqref{eq:abjm-superpotential}; closure
of the enhanced supersymmetry fixes its coefficient to
\[
  h_{\rm ABJM}=\frac{2\pi}{k}
\]
in the component convention of this chapter.  Deformations that keep only
\(\mathcal N=2\), change the flavor symmetry, change the line lattice, or
alter background contact terms are different QFT specifications and are not counted as
tangent vectors to this standard ABJM conformal class.
\end{hypothesis}

\begin{remark}[Standard ABJM has no continuous conformal coupling]
\label{rem:standard-abjm-zero-dimensional-conformal-locus}
Assume Hypothesis~\ref{hyp:standard-abjm-conformal-class}.  At fixed
\((N,k)\) the local conformal-manifold germ of the standard ABJM specification is
zero-dimensional.  The standard ABJM theories form a discrete family indexed
by the integer level \(k\), with \(k\) and \(-k\) related when one includes
orientation reversal and exchange of the two gauge factors as described above.

The Chern--Simons levels are not coordinates on a real or complex manifold.
Gauge invariance of \(\exp(\ii S_{\rm CS})\) under large gauge
transformations restricts the levels to an integral lattice.  Hence an
infinitesimal deformation preserving the same gauge specification has
\[
  \delta k=0 .
\]
The opposite-level condition in the standard ABJM specification also fixes the
second level to be \(-k\), so there is no independent continuous level
direction.

The quartic superpotential is power-counting marginal in an
\(\mathcal N=2\) description: the chiral fields have protected
\(\Delta=R=\frac12\) at the ABJM point, so the quartic monomial has
dimension \(2\), and the \(d^2\theta\) measure has dimension \(1\) in three
dimensions.  Power-counting marginality, however, is only a necessary test.
Inside the standard \(\mathcal N=6\) ABJM class the same coefficient is
fixed by the supersymmetry-completion condition
\[
  h=h_{\rm ABJM}=\frac{2\pi}{k}.
\]
Therefore
\[
  \delta h=-\frac{2\pi}{k^2}\,\delta k=0
\]
for a tangent vector that preserves the standard ABJM specification.

The remaining familiar Lagrangian coordinates do not give exactly marginal
directions in this class.  Real masses and FI coordinates have engineering
dimension \(1\) in three dimensions and break part of the conformal or flavor
specification.  A Yang--Mills kinetic term, if introduced as a regulator or UV
completion coordinate, carries a dimensionful coupling
\[
  [g_{\rm YM}^2]=1
\]
and is not a scalar-primary deformation of dimension \(3\) of the conformal
theory.  Background Chern--Simons contact terms are quantized scheme data,
not continuous local-operator deformations.  Thus every allowed tangent
vector in the fixed standard ABJM specification is redundant or zero.
\end{remark}

This remark is deliberately narrower than a classification theorem for
all three-dimensional \(\mathcal N=2\) Chern--Simons--matter theories with
the same gauge group.  If the \(SU(4)\) symmetry, the \(\mathcal N=6\)
completion, or the standard line and contact-term data are relaxed, quartic
marginal candidates must be analyzed by their own beta functions, current
moment maps, and monopole-sector constraints.  They are not continuous
coordinates on the standard ABJM conformal locus just defined.

\section{ABJM Classical Moduli Space}
\label{sec:abjm-classical-moduli-space}

The vacuum equations of the \(G_{N,k}\) theory are not those of an ordinary
Yang--Mills quotient: the Chern--Simons auxiliary fields identify moment-map
coordinates with vector-multiplet scalars.  A useful branch is the commuting
branch, where the four complex scalar matrices are simultaneously diagonal
after gauge transformations:
\[
  C_I=\operatorname{diag}(c_{I,1},\ldots,c_{I,N}),\qquad I=1,\ldots,4 .
\]
For generic diagonal eigenvalues the residual gauge group is the permutation
group \(S_N\), together with abelian gauge transformations acting on each
eigenvalue.

\paragraph{Rank-one abelian quotient.}
For \(N=1\), set \(C=(C_1,\ldots,C_4)\).  The classical ABJM branch is
\[
  \mathcal M_{1,k}^{\rm cl}=\C^4/\mathbb Z_k,
\]
where the generator of \(\mathbb Z_k\) acts by simultaneous phase rotation
\[
  (C_1,\ldots,C_4)\longmapsto
  \bigl(\ee^{2\pi\ii/k}C_1,\ldots,\ee^{2\pi\ii/k}C_4\bigr).
\]

For \(U(1)_k\times U(1)_{-k}\), every matter scalar has charge \(+1\) under
the anti-diagonal gauge parameter and charge \(0\) under the diagonal gauge
parameter.  Put
\[
  A_D=A+\widehat A,\qquad B=A-\widehat A .
\]
On a closed three-manifold the abelian Chern--Simons term is, up to the
declared boundary convention,
\begin{equation}
  \frac{k}{4\pi}\int(A\wedge dA-\widehat A\wedge d\widehat A)
  =
  \frac{k}{4\pi}\int B\wedge dA_D .
  \label{eq:abjm-abelian-bf-term}
\end{equation}
Thus the diagonal gauge field acts as a Lagrange multiplier for the
anti-diagonal flux, while the anti-diagonal gauge field is the one seen by
the matter.  Dualizing the flat diagonal connection gives a periodic scalar.
Large shifts of that scalar leave a residual anti-diagonal gauge
transformation of order \(k\).  This residual transformation multiplies every
\(C_I\) by \(\ee^{2\pi\ii/k}\).  Therefore the continuous anti-diagonal
phase quotient is accompanied by the finite residual quotient
\(\mathbb Z_k\), giving \(\C^4/\mathbb Z_k\).

\begin{quotedtheorem}[ABJM commuting-branch moduli space]
\label{qt:abjm-commuting-branch-moduli}
For the \(U(N)_k\times U(N)_{-k}\) ABJM theory, the branch obtained from
simultaneously diagonalizable scalar vacuum expectation values has classical
moduli space
\[
  \mathcal M_{N,k}^{\rm cl}
  =
  \bigl((\C^4/\mathbb Z_k)^N\bigr)/S_N .
\]
For \(k=1,2\) the interacting theory has enhanced \(\mathcal N=8\)
supersymmetry; for \(|k|>2\) the standard interacting theory has
\(\mathcal N=6\) supersymmetry.  The enhancement statement uses monopole
operators and requires data beyond the polynomial superpotential.
\end{quotedtheorem}

The theorem is quoted here because the full statement identifies quantum
operator sectors, monopole operators, and the global form of the gauge group.
The finite quotient in the rank-one calculation above is the intrinsic
gauge-theory convention that fixes the \(k\)-dependent orbifold action used
later in localization and protected-sector comparisons.

\section{ABJM Localization Specialization}
\label{sec:abjm-localization-specialization}

The \(S^3\) localization formula of
Chapter~\ref{ch:susy-localization-compact-manifolds} specializes to ABJM
when all four chiral multiplets have \(R\)-charge \(1/2\) and zero real mass.
For Cartan variables
\[
  \lambda_i,\mu_j\in\mathbb R,\qquad i,j=1,\ldots,N,
\]
associated respectively to the two \(U(N)\) factors, the matrix integral is
\begin{equation}
\begin{aligned}
  Z^{\rm ABJM}_{S^3}(N,k)
  &=
  \frac1{(N!)^2}
  \int_{\mathbb R^{2N}}
  \prod_{i=1}^N d\lambda_i\,d\mu_i\,
  \exp\!\left[
    \ii\pi k\sum_i(\lambda_i^2-\mu_i^2)
  \right] \\
  &\quad\times
  \frac{
    \prod_{i<j}4\sinh^2\!\bigl(\pi(\lambda_i-\lambda_j)\bigr)
    \prod_{i<j}4\sinh^2\!\bigl(\pi(\mu_i-\mu_j)\bigr)
  }{
    \prod_{i,j}4\cosh^2\!\bigl(\pi(\lambda_i-\mu_j)\bigr)
  } .
\end{aligned}
  \label{eq:abjm-s3-matrix-model}
\end{equation}
The numerator is the product of the two vector-multiplet determinants.  The
denominator is the product of two conjugate bifundamental chiral pairs:
each pair contributes
\[
  \exp\!\left[
    \ell\!\left(\frac12+\ii(\lambda_i-\mu_j)\right)
    +
    \ell\!\left(\frac12-\ii(\lambda_i-\mu_j)\right)
  \right]
  =
  \frac1{2\cosh\!\bigl(\pi(\lambda_i-\mu_j)\bigr)} .
\]
The two pairs give \(1/(4\cosh^2(\pi(\lambda_i-\mu_j)))\) for each ordered
pair \((i,j)\).  The oscillatory contour, framing phase, and contact terms
are part of the localization input.  Thus
\eqref{eq:abjm-s3-matrix-model} is an exact protected observable once those
inputs are supplied.  A microscopic construction of the continuum theory
requires independent definition of the underlying three-dimensional QFT.

\section{Sphere Free Energy and the F-Theorem}
\label{sec:three-d-sphere-free-energy-f-theorem}

Three-dimensional conformal field theories have a distinguished finite
observable on the round sphere.  It is useful in supersymmetric
Chern--Simons--matter theory because localization often reduces it to a
finite-dimensional matrix integral, but its definition is not a localization
artifact.

\begin{definition}[Renormalized three-sphere free energy]
\label{def:renormalized-s3-free-energy}
Let \(\mathcal T\) be a reflection-positive three-dimensional Euclidean QFT
with a specified regulator, background metric coupling, and local
counterterm scheme.  Its round-sphere partition function at radius \(r\) is
denoted \(Z_{S^3_r}(\mathcal T)\).  At a conformal fixed point, after
removing power divergences by the declared local counterterms, the
renormalized sphere free energy is
\begin{equation}
  F(\mathcal T)
  =
  -\log\left|Z_{S^3}(\mathcal T)\right|.
  \label{eq:s3-free-energy-definition}
\end{equation}
The absolute value removes pure phases produced by framing and background
Chern--Simons contact terms.  It does not remove the obligation to declare
the real local counterterm scheme used to define the finite partition
function.
\end{definition}

In a supersymmetric theory the same definition is evaluated by the localized
matrix integral only after the \(S^3\) localization input of
Chapter~\ref{ch:susy-localization-compact-manifolds} has been supplied:
Cartan contour, parity-anomaly regulator, framing phase, real-mass and FI
coordinates, and background Chern--Simons counterterms.  For ABJM,
\begin{equation}
  F_{\rm ABJM}(N,k)
  =
  -\log\left|Z^{\rm ABJM}_{S^3}(N,k)\right|
  \label{eq:abjm-sphere-free-energy}
\end{equation}
with \(Z^{\rm ABJM}_{S^3}\) defined by
\eqref{eq:abjm-s3-matrix-model}.  This is a protected observable of the
declared QFT specification; changing the line-operator lattice or contact terms can
change phases and, in some schemes, finite background terms.

\begin{quotedtheorem}[\(F\)-theorem]
\label{qt:continuum-f-theorem}
Let a unitary, Lorentz-invariant three-dimensional QFT define an RG flow
between ultraviolet and infrared conformal fixed points, and assume the
stress tensor, vacuum state, ball entanglement entropy, and continuum limit
are controlled so that strong subadditivity and the Casini--Huerta
renormalized-entanglement construction apply.  Define
\begin{equation}
  \mathcal F(R)=R\,\frac{dS(R)}{dR}-S(R),
  \label{eq:renormalized-entanglement-f}
\end{equation}
where \(S(R)\) is the vacuum entanglement entropy of a disk of radius \(R\).
Then \(\mathcal F(R)\) is monotone nonincreasing as \(R\) increases.  At a
conformal fixed point, the conformal map from the causal development of the
disk to the hyperbolic thermal problem identifies \(\mathcal F\) with the
renormalized sphere free energy \(F\).  Consequently a nontrivial flow
satisfying these hypotheses obeys
\begin{equation}
  F_{\rm UV}>F_{\rm IR},
  \label{eq:f-theorem-inequality}
\end{equation}
with equality only for a trivial flow modulo decoupled topological/contact
data.
\end{quotedtheorem}

\begin{remark}[The theorem is not a matrix-model identity]
\label{rem:f-theorem-not-matrix-model-identity}
The monotonicity theorem is a continuum theorem about unitary RG flows.  A
localized \(S^3\) matrix model computes \(F\) at a supersymmetric fixed point
when its localization input is valid; it does not by itself prove
monotonicity for flows between two fixed points.  Conversely, the
\(F\)-theorem applies to nonsupersymmetric flows for which no localization
formula exists.
\end{remark}

\begin{remark}[Consequences for supersymmetric fixed-point comparisons]
\label{rem:f-theorem-susy-comparison}
Suppose \(\mathcal T_{\rm UV}\to\mathcal T_{\rm IR}\) is a three-dimensional
unitary RG flow satisfying the hypotheses of
Quoted Theorem~\ref{qt:continuum-f-theorem}, and suppose both endpoints have
supersymmetric localization data whose finite parts compute the
renormalized \(S^3\) partition functions in the scheme of
Definition~\ref{def:renormalized-s3-free-energy}.  Then the localized
numbers obey
\[
  -\log|Z_{S^3}^{\rm loc}(\mathcal T_{\rm UV})|
  >
  -\log|Z_{S^3}^{\rm loc}(\mathcal T_{\rm IR})|
\]
for a nontrivial flow.  For a product of decoupled fixed points
\(\mathcal T_1\otimes\mathcal T_2\), the sphere free energy is additive:
\begin{equation}
  F(\mathcal T_1\otimes\mathcal T_2)=F(\mathcal T_1)+F(\mathcal T_2).
  \label{eq:s3-free-energy-additivity}
\end{equation}

The first assertion is the inequality
\eqref{eq:f-theorem-inequality} after identifying the endpoint \(F\)'s with
the localized finite parts.  This identification is an endpoint statement:
it uses the localization input at the fixed point and does not require a
localized description of the entire RG trajectory.

For the product statement, the Hilbert space, operator algebra, and
Euclidean functional integral factorize.  With product regulators and
counterterm schemes,
\[
  Z_{S^3}(\mathcal T_1\otimes\mathcal T_2)
  =
  Z_{S^3}(\mathcal T_1)Z_{S^3}(\mathcal T_2).
\]
Taking \(-\log|\cdot|\) gives
\eqref{eq:s3-free-energy-additivity}.  Possible background contact terms
also add because they are local functionals of the external fields.
\end{remark}

\begin{example}[Free chiral multiplet and a massive flow]
\label{ex:free-chiral-f-value}
For a free \(\mathcal N=2\) chiral multiplet of \(R\)-charge \(1/2\) and no
background mass, the round-sphere determinant convention of
Chapter~\ref{ch:susy-localization-compact-manifolds} gives
\begin{equation}
  Z_{\rm chiral}=\exp\ell\!\left(\frac12\right)=2^{-1/2},
  \qquad
  F_{\rm chiral}=\frac12\log2 .
  \label{eq:free-chiral-f-value}
\end{equation}
Giving this chiral multiplet a supersymmetric real mass and flowing to
energies much smaller than \(|m|\) leaves the empty gapped theory, whose
non-topological sphere free energy is \(0\) in the same normalization.  The
flow therefore has
\[
  F_{\rm UV}=\frac12\log2>0=F_{\rm IR},
\]
as required by the \(F\)-theorem.  This example is simple, but it fixes the
normalization used when more complicated localized matrix integrals are
compared.
\end{example}

\paragraph{Rank-one ABJM sphere integral.}
With the oscillatory contour and framing convention implicit in
\eqref{eq:abjm-s3-matrix-model}, the \(N=1\) ABJM matrix integral evaluates
distributionally to
\begin{equation}
  Z^{\rm ABJM}_{S^3}(1,k)=\frac{1}{4|k|},
  \qquad
  F_{\rm ABJM}(1,k)=\log(4|k|).
  \label{eq:rank-one-abjm-s3-value}
\end{equation}

For \(N=1\), the vector Vandermonde factors are absent and
\eqref{eq:abjm-s3-matrix-model} becomes
\[
  Z
  =
  \int_{\mathbb R^2}d\lambda\,d\mu\,
  \frac{\exp\{\ii\pi k(\lambda^2-\mu^2)\}}
       {4\cosh^2(\pi(\lambda-\mu))}.
\]
Set
\[
  x=\lambda-\mu,\qquad y=\lambda+\mu .
\]
Then \(d\lambda\,d\mu=\frac12\,dx\,dy\) and
\(\lambda^2-\mu^2=xy\).  In the oscillatory distribution sense,
\[
  \int_{-\infty}^{\infty}dy\,\ee^{\ii\pi kxy}
  =
  2\,\delta(kx)
  =
  \frac{2}{|k|}\delta(x).
\]
Therefore
\[
  Z
  =
  \frac12\int dx\,
  \frac{1}{4\cosh^2(\pi x)}
  \frac{2}{|k|}\delta(x)
  =
  \frac{1}{4|k|}.
\]
Taking the absolute value and then \(-\log\) gives the displayed
free energy.

\section{ABJM Fermi-Gas Form and Airy Asymptotics}
\label{sec:abjm-fermi-gas-airy}

The ABJM sphere integral has a second exact representation which is useful
because it separates three logically different statements:
\begin{enumerate}
\item the finite-dimensional localization integral
\eqref{eq:abjm-s3-matrix-model};
\item an algebraic rewriting of that integral as the canonical partition
function of \(N\) noninteracting fermions with a one-particle density
operator;
\item the large-\(N\) Airy behavior that follows when the grand potential of
that one-particle operator has a cubic perturbative part.
\end{enumerate}
Only the second item is a change of variables in the localized integral.  It
does not define the continuum ABJM theory and it does not replace the
localization hypotheses.  Throughout this section take \(k>0\); negative
level is obtained by orientation reversal and complex conjugation of the
Chern--Simons phase.

Let \(\hbar=2\pi k\), let \(q\) denote multiplication on
\(L^2(\mathbb R,dq)\), and let \(p=-\ii\hbar\,d/dq\).  Define
\[
  u(q)=\frac1{\sqrt{2\cosh(q/2)}},
  \qquad
  v(p)=\frac1{2\cosh(p/2)}.
\]
The ABJM one-particle density operator is
\begin{equation}
  \rho_k
  =
  u(q)\,v(p)\,u(q).
  \label{eq:abjm-density-operator}
\end{equation}
Equivalently, in the \(q\)-representation,
\begin{equation}
  \rho_k(q,q')
  =
  \frac1{2\pi k}\,
  \frac1{\sqrt{2\cosh(q/2)}}\,
  \frac1{2\cosh\!\left(\frac{q-q'}{2k}\right)}\,
  \frac1{\sqrt{2\cosh(q'/2)}} .
  \label{eq:abjm-fermi-gas-kernel}
\end{equation}

\begin{lemma}[Trace-class normalization of the ABJM kernel]
\label{lem:abjm-kernel-trace-class}
The operator \(\rho_k\) in \eqref{eq:abjm-density-operator} is positive and
trace class.  Its first trace is
\begin{equation}
  \operatorname{Tr}_{\mathcal H}\rho_k=\frac1{4k}.
  \label{eq:abjm-rho-first-trace}
\end{equation}
\end{lemma}

\begin{proof}
Use the unitary Fourier transform normalized by
\[
  \langle q|p\rangle=(2\pi\hbar)^{-1/2}\ee^{\ii qp/\hbar}.
\]
The multiplication operator \(v(p)\) is positive.  Hence
\[
  \rho_k
  =
  \left(v(p)^{1/2}u(q)\right)^*
  \left(v(p)^{1/2}u(q)\right)
\]
is positive.  The operator \(v(p)^{1/2}u(q)\) is Hilbert--Schmidt, because
its \(p,q\) kernel has squared norm
\[
  \int_{\mathbb R^2}\frac{dp\,dq}{2\pi\hbar}\,
  \frac1{2\cosh(p/2)}\,\frac1{2\cosh(q/2)}
  <\infty .
\]
Therefore \(\rho_k=T^*T\) with \(T\) Hilbert--Schmidt, so \(\rho_k\) is
trace class.

The Fourier integral
\[
  \int_{-\infty}^{\infty}
  \frac{dp}{2\pi\hbar}\,
  \frac{\ee^{\ii p(q-q')/\hbar}}{2\cosh(p/2)}
  =
  \frac1{2\pi k}\,
  \frac1{2\cosh\!\left(\frac{q-q'}{2k}\right)}
\]
gives \eqref{eq:abjm-fermi-gas-kernel}.  On the diagonal,
\[
  \rho_k(q,q)
  =
  \frac1{2\pi k}\,
  \frac1{4\cosh(q/2)} .
\]
Since
\[
  \int_{-\infty}^{\infty}\frac{dq}{\cosh(q/2)}=2\pi,
\]
the trace is \(1/(4k)\).
\end{proof}

\begin{quotedtheorem}[Finite ABJM Fermi-gas identity]
\label{qt:abjm-fermi-gas-representation}
For \(k>0\), the ABJM localization integral
\eqref{eq:abjm-s3-matrix-model} is equal to
\begin{equation}
  Z^{\rm ABJM}_{S^3}(N,k)
  =
  \frac1{N!}\int_{\mathbb R^N}\prod_{i=1}^N dq_i\,
  \det_{1\le i,j\le N}\rho_k(q_i,q_j).
  \label{eq:abjm-canonical-fermi-gas}
\end{equation}
\end{quotedtheorem}

\begin{remark}[What enters the finite identity]
\label{rem:abjm-fermi-gas-identity-ingredients}
The identity is finite-dimensional algebra inside the localized matrix
model, not a holographic or continuum-existence statement.  Its input is the
change of variables
\[
  x_i=2\pi\lambda_i,\qquad y_i=2\pi\mu_i .
\]
The main determinant step is the hyperbolic Cauchy identity
\begin{equation}
  \frac{\prod_{i<j}2\sinh\!\left(\frac{x_i-x_j}{2}\right)
        \prod_{i<j}2\sinh\!\left(\frac{y_i-y_j}{2}\right)}
       {\prod_{i,j}2\cosh\!\left(\frac{x_i-y_j}{2}\right)}
  =
  \det_{i,j}\frac1{2\cosh\!\left(\frac{x_i-y_j}{2}\right)} .
  \label{eq:hyperbolic-cauchy-abjm}
\end{equation}
It follows from the ordinary Cauchy identity after setting
\(X_i=\ee^{x_i}\), \(Y_j=\ee^{y_j}\) and extracting monomials from rows and
columns.  Squaring \eqref{eq:hyperbolic-cauchy-abjm} rewrites the
Vandermonde ratio in \eqref{eq:abjm-s3-matrix-model} as a product of
determinants.  The remaining step is a Gaussian Fourier transform plus a
unitary phase conjugation.  Phase conjugation leaves every Fredholm minor
unchanged:
\[
  \det\!\left[
    \ee^{\ii k q_i^2/4\pi}\rho_k(q_i,q_j)
    \ee^{-\ii k q_j^2/4\pi}
  \right]_{i,j=1}^N
  =
  \det\!\left[\rho_k(q_i,q_j)\right]_{i,j=1}^N .
\]
The Fourier normalization is the one displayed in the proof of
Lemma~\ref{lem:abjm-kernel-trace-class}.  For \(N=1\),
\eqref{eq:abjm-canonical-fermi-gas} gives
\(\operatorname{Tr}\rho_k=1/(4k)\), agreeing with
the rank-one sphere-integral calculation above.
\end{remark}

\begin{definition}[ABJM grand partition function]
\label{def:abjm-grand-partition-function}
The grand partition function attached to the ABJM Fermi-gas operator is
\begin{equation}
  \Xi_k(\kappa)
  =
  \det_{\mathcal H}(1+\kappa\rho_k),
  \qquad
  \kappa=\ee^\mu ,
  \label{eq:abjm-grand-partition-function}
\end{equation}
and its logarithm
\begin{equation}
  J_k(\mu)=\log\Xi_k(\ee^\mu)
  \label{eq:abjm-grand-potential}
\end{equation}
is the grand potential.
\end{definition}

\begin{lemma}[Fredholm expansion and canonical coefficients]
\label{lem:fredholm-canonical-coefficients}
Let \(\rho\) be a trace-class operator on \(L^2(\mathbb R,dq)\) with
integral kernel \(\rho(q,q')\).  Then
\begin{equation}
  \det(1+\kappa\rho)
  =
  \sum_{N=0}^{\infty}\kappa^N
  \frac1{N!}\int_{\mathbb R^N}\prod_{i=1}^N dq_i\,
  \det_{i,j}\rho(q_i,q_j).
  \label{eq:fredholm-minor-expansion}
\end{equation}
Consequently the coefficient of \(\kappa^N\) in
\(\Xi_k(\kappa)\) is the ABJM sphere partition function
\(Z^{\rm ABJM}_{S^3}(N,k)\).
\end{lemma}

\begin{proof}
For finite-rank \(\rho=\sum_{a=1}^r s_a\,|\psi_a\rangle\langle\phi_a|\),
the identity is the Cauchy--Binet expansion of the finite determinant
\(\det(\delta_{ab}+\kappa s_a\langle\phi_a,\psi_b\rangle)\).  Approximate a
trace-class operator in trace norm by finite-rank operators.  The Fredholm
determinant is continuous in trace norm, and the Fredholm minors converge in
the same expansion because the exterior powers satisfy
\(\|\wedge^N\rho\|_1\le \|\rho\|_1^N/N!\).  This gives
\eqref{eq:fredholm-minor-expansion}.  Applying
Quoted Theorem~\ref{qt:abjm-fermi-gas-representation} identifies the
coefficients with the localized ABJM integrals.
\end{proof}

\paragraph{Leading phase-space coefficient.}
The leading Weyl coefficient of the ABJM Fermi gas is
\begin{equation}
  C_k=\frac{2}{\pi^2 k}.
  \label{eq:abjm-Ck-coefficient}
\end{equation}
Equivalently, the leading large-\(\mu\) contribution to the grand potential
is
\begin{equation}
  J_k(\mu)=\frac{C_k}{3}\mu^3+\text{lower powers of }\mu
  \qquad (\mu\to+\infty)
  \label{eq:abjm-leading-grand-potential}
\end{equation}
at fixed \(k\).

The classical symbol of \(\rho_k^{-1}\) is
\[
  \exp H(q,p),
  \qquad
  H(q,p)
  =
  \log\!\left(2\cosh\frac q2\right)
  +
  \log\!\left(2\cosh\frac p2\right).
\]
For large \(E\), the inequality \(H(q,p)\le E\) has the same leading area
as
\[
  \frac{|q|+|p|}{2}\le E,
\]
namely a diamond of area \(8E^2\).  The phase-space cell is
\[
  2\pi\hbar=4\pi^2 k .
\]
Thus the integrated one-particle density has leading term
\[
  n(E)
  =
  \frac{8E^2}{4\pi^2 k}+O(E)
  =
  \frac{2}{\pi^2 k}E^2+O(E).
\]
At zero temperature the leading grand potential is
\[
  J_k(\mu)
  =
  \int_0^\mu n(E)\,dE+\text{lower powers of }\mu
  =
  \frac{2}{3\pi^2 k}\mu^3+\text{lower powers},
\]
which gives \eqref{eq:abjm-Ck-coefficient}.

\begin{quotedtheorem}[Perturbative ABJM grand potential]
\label{qt:abjm-perturbative-grand-potential}
For the trace-class operator \(\rho_k\) above, the part of \(J_k(\mu)\)
which is polynomial in \(\mu\) in the large-\(\mu\) spectral expansion is
\begin{equation}
  J_k^{\rm pert}(\mu)
  =
  \frac{C_k}{3}\mu^3+B_k\mu+A_k,
  \qquad
  C_k=\frac{2}{\pi^2 k},
  \qquad
  B_k=\frac{k}{24}+\frac1{3k}.
  \label{eq:abjm-perturbative-grand-potential}
\end{equation}
The remaining terms in \(J_k(\mu)\) are exponentially small in \(\mu\)
within the spectral expansion and encode nonperturbative corrections to the
canonical large-\(N\) answer.  The coefficient \(A_k\) is independent of
\(\mu\) and depends on the normalization of the exact spectral determinant.
\end{quotedtheorem}

The inverse Laplace transform of a cubic perturbative grand potential is an
Airy function.  Assume
\begin{equation}
  J^{\rm pert}(\mu)=\frac C3\mu^3+B\mu+A,
  \qquad C>0.
  \label{eq:cubic-grand-potential-general}
\end{equation}
Define the perturbative canonical coefficient by the inverse Laplace
transform
\begin{equation}
  Z_N^{\rm pert}
  =
  \int_{\mathcal C}\frac{d\mu}{2\pi\ii}\,
  \exp\!\left(J^{\rm pert}(\mu)-N\mu\right),
  \label{eq:perturbative-canonical-inverse-laplace}
\end{equation}
where \(\mathcal C\) is the standard Airy contour.  Then
\begin{equation}
  Z_N^{\rm pert}
  =
  \ee^A\,C^{-1/3}\,
  \operatorname{Ai}\!\left(C^{-1/3}(N-B)\right).
  \label{eq:airy-transform-result}
\end{equation}
For ABJM this gives
\begin{equation}
  Z_{S^3}^{\rm ABJM,pert}(N,k)
  =
  \ee^{A_k}\,
  C_k^{-1/3}\,
  \operatorname{Ai}\!\left(C_k^{-1/3}(N-B_k)\right),
  \label{eq:abjm-perturbative-airy}
\end{equation}
with \(C_k,B_k\) as in
\eqref{eq:abjm-perturbative-grand-potential}.
Set \(t=C^{1/3}\mu\).  Then \(d\mu=C^{-1/3}dt\) and
\[
  \frac C3\mu^3-(N-B)\mu
  =
  \frac{t^3}{3}
  -
  C^{-1/3}(N-B)t .
\]
Therefore
\[
  Z_N^{\rm pert}
  =
  \ee^A C^{-1/3}
  \int_{\mathcal C_A}\frac{dt}{2\pi\ii}\,
  \exp\!\left(\frac{t^3}{3}
  -
  C^{-1/3}(N-B)t\right).
\]
The last integral is the contour definition of the Airy function
\(\operatorname{Ai}(z)\) with \(z=C^{-1/3}(N-B)\).  This proves
\eqref{eq:airy-transform-result}.  Substituting the ABJM coefficients gives
\eqref{eq:abjm-perturbative-airy}.

\begin{remark}[Status of the ABJM Airy formula]
\label{rem:abjm-airy-status}
The Airy function in \eqref{eq:abjm-perturbative-airy} is an exact
consequence of the cubic perturbative grand potential; it is not by itself a
statement about the full finite-\(N\) sphere partition function.  The full
Fredholm determinant \(\Xi_k(\kappa)\) also contains exponentially small
spectral corrections, and their inverse Laplace transforms correct the
Airy expression.  Thus the logically complete chain is:
localization matrix integral, Fermi-gas Fredholm determinant, spectral
large-\(\mu\) expansion, Airy transform of the cubic part, and then
nonperturbative corrections.  Confusing the last two steps is a common
source of overstated ``exact large-\(N\)'' claims.
\end{remark}

\section{Three-Dimensional Dynamics Ledger}

A complete \(3D\) supersymmetric Chern--Simons--matter example must specify:
\[
\begin{gathered}
  G,\quad k,\quad R,\quad W,\quad \zeta,\quad m_i,\\
  \text{spin structure and framing data},\quad
  \text{parity-anomaly regulator},\\
  \text{monopole-operator dressing rules},\quad
  \text{boundary and global-form choices}.
\end{gathered}
\]
For ABJM this list becomes
\[
\begin{gathered}
  U(N)_k\times U(N)_{-k},\quad
  A_a,B_{\dot a},\quad W_{\rm ABJM},\\
  \text{the discrete level } k \text{ and the zero-dimensional standard
  conformal locus},\\
  \text{the chosen global form and line-operator lattice},\\
  \text{the monopole sectors implementing the }SU(4)\text{ symmetry and
  possible } \mathcal N=8 \text{ enhancement},\\
  \text{and the }S^3\text{ contour, framing, and contact-term convention}.
\end{gathered}
\]
Duality proposals compare these full data, including contact terms and
background Chern--Simons counterterms.

\begin{openproblem}[Three-dimensional supersymmetric continuum schemes]
\label{op:three-dimensional-susy-cs-matter-continuum}
Construct \(3D\) supersymmetric Chern--Simons--matter theories with a
regulator preserving the stated gauge and supersymmetry Ward identities,
including parity-anomaly choices, monopole-operator definitions, boundary
conditions, contact-term coordinates, and global-form data.
\end{openproblem}
