\chapter{Four-Dimensional N=1 Gauge Dynamics}
\label{ch:susy-four-dimensional-n1-gauge-dynamics}

\ConventionsLorentzian

This chapter develops the first layer of four-dimensional \(\mathcal N=1\)
gauge dynamics in Wilsonian coordinates.  The conclusions are conditional on
the supersymmetric Wilsonian scheme and anomaly calculation fixed in
Chapters~\ref{ch:susy-wilsonian-schemes-exact-dynamics} and
\ref{ch:susy-nonrenormalization-holomorphy}.

\section{SQCD Data}

Let \(G=SU(N_c)\).  Supersymmetric QCD with \(N_f\) flavors has chiral
superfields
\[
  Q^i\in \mathbf N_c,\qquad
  \widetilde Q_j\in \overline{\mathbf N}_c,
  \qquad i,j=1,\ldots,N_f .
\]
The vector multiplet has field strength \(W_\alpha\), and the Wilsonian gauge
coordinate is
\[
  \frac{1}{16\pi\ii}\int d^4x\,d^2\theta\,
  \tau\,\operatorname{tr}(W^\alpha W_\alpha)+\text{c.c.}
\]
with the trace convention inherited from Chapter~\ref{ch:susy-gauge-theory}.
The holomorphic scale is the invariant coordinate
\[
  \Lambda_h^{b_0}
  =
  \mu^{b_0}\exp(2\pi\ii\tau(\mu)),
  \qquad
  b_0=3N_c-N_f .
\]
This equation defines \(\Lambda_h\) after a branch convention has been chosen;
only \(\Lambda_h^{b_0}\) is invariant under the periodicity of \(\theta\).

\section{SQCD NSVZ Coordinate Audit}
\label{sec:sqcd-nsvz-coordinate-audit}

The holomorphic scale is not the same coordinate as the canonical gauge
coupling used in particle normalizations.  This distinction is a coordinate
statement, not a change of the underlying theory.  Write
\[
  X_h(\mu)=\frac{8\pi^2}{g_h(\mu)^2},\qquad
  X_c(\mu)=\frac{8\pi^2}{g(\mu)^2},
\]
where \(g_h\) is the holomorphic Wilsonian coupling and \(g\) is the
canonical coupling defined by the component kinetic term of
Chapter~\ref{ch:susy-gauge-theory}.  Let \(Z_Q\) and
\(Z_{\widetilde Q}\) be the matter wave-function coordinates, with
\[
  \gamma_Q(\mu)=\mu\frac{d}{d\mu}\log Z_Q(\mu),
  \qquad
  \gamma_{\widetilde Q}(\mu)
  =
  \mu\frac{d}{d\mu}\log Z_{\widetilde Q}(\mu).
\]

\begin{hypothesis}[SQCD NSVZ coordinate specialization]
\label{hyp:sqcd-nsvz-coordinate-specialization}
The SQCD beta-function identity below is stated in a perturbative
supersymmetric Wilsonian scheme satisfying the hypotheses of
Propositions~\ref{prop:holomorphic-gauge-coupling-one-loop-exact} and
\ref{prop:nsvz-coordinate-relation-derived}: in particular the \(q_h^0\)
closure input for the holomorphic gauge coordinate is part of the scheme
datum alongside theta periodicity.  The
flavor-symmetric submanifold is imposed by taking
\[
  Z_Q=Z_{\widetilde Q},\qquad
  \gamma_Q=\gamma_{\widetilde Q}.
\]
The trace convention is the half-trace convention of this chapter:
\[
  C_2(SU(N_c))=N_c,\qquad
  T(\mathbf N_c)=T(\overline{\mathbf N}_c)=\frac12 .
\]
Finite changes of canonical coupling coordinate are allowed, but then the
rational NSVZ denominator changes by the induced reparametrization.  Thus
the pole of the displayed denominator is a feature of this coordinate chart;
the invariant tests in later sections are the numerator condition, anomaly
matching, holomorphic scale matching, and protected chiral data.
\end{hypothesis}

\paragraph{SQCD specialization of the NSVZ coordinate relation.}
Assume Hypothesis~\ref{hyp:sqcd-nsvz-coordinate-specialization}.  In the
flavor-symmetric SQCD coordinate chart,
\begin{equation}
  X_h
  =
  X_c+N_c\log g^2-N_f\log Z_Q+\kappa ,
  \label{eq:sqcd-holomorphic-canonical-coordinate-relation}
\end{equation}
where \(\kappa\) is independent of \(\mu\).  Consequently
\begin{equation}
  \beta_g
  =
  -\frac{g^3}{16\pi^2}\,
  \frac{
    3N_c-N_f(1-\gamma_Q)
  }{
    1-\frac{N_cg^2}{8\pi^2}
  } .
  \label{eq:sqcd-nsvz-beta-specialized}
\end{equation}

The general relation of
Proposition~\ref{prop:nsvz-coordinate-relation-derived} gives
\[
  X_h
  =
  X_c+C_2(SU(N_c))\log g^2
  -
  \sum_i T(R_i)\log Z_i+\kappa .
\]
There are \(N_f\) chiral multiplets \(Q^i\) in \(\mathbf N_c\) and \(N_f\)
chiral multiplets \(\widetilde Q_j\) in \(\overline{\mathbf N}_c\).  Hence
\[
  \sum_iT(R_i)\log Z_i
  =
  \frac{N_f}{2}\log Z_Q+\frac{N_f}{2}\log Z_{\widetilde Q}.
\]
On the flavor-symmetric submanifold this is \(N_f\log Z_Q\), and
\(C_2(SU(N_c))=N_c\).  This proves
\eqref{eq:sqcd-holomorphic-canonical-coordinate-relation}.

By Proposition~\ref{prop:holomorphic-gauge-coupling-one-loop-exact}, the
holomorphic Wilsonian coordinate has one-loop derivative
\[
  \mu\frac{dX_h}{d\mu}=3N_c-N_f .
\]
Differentiating \eqref{eq:sqcd-holomorphic-canonical-coordinate-relation}
therefore gives
\[
  \left(
    -\frac{16\pi^2}{g^3}+\frac{2N_c}{g}
  \right)\beta_g
  -N_f\gamma_Q
  =
  3N_c-N_f .
\]
Moving the anomalous-dimension term to the right hand side yields the
NSVZ numerator \(3N_c-N_f(1-\gamma_Q)\).  Solving the displayed linear
equation for \(\beta_g=\mu\,dg/d\mu\) gives
\eqref{eq:sqcd-nsvz-beta-specialized}.

Equivalently, inserting
\[
  C_2(SU(N_c))=N_c,\qquad
  T(\mathbf N_c)=T(\overline{\mathbf N}_c)=\frac12
\]
directly into the general NSVZ coordinate formula of
Chapter~\ref{ch:susy-nonrenormalization-holomorphy} gives
\[
  \beta_g
  =
  -\frac{g^3}{16\pi^2}\,
  \frac{
    3C_2(SU(N_c))
    -
    N_f\bigl(T(\mathbf N_c)+T(\overline{\mathbf N}_c)\bigr)
    (1-\gamma_Q)
  }{
    1-\frac{C_2(SU(N_c))g^2}{8\pi^2}
  },
\]
whose numerator is
\[
  3N_c-N_f(1-\gamma_Q).
\]
The proof above displays why the same numerator arises from the
holomorphic-canonical coordinate relation: the two factors of \(1/2\) from
the fundamental and antifundamental indices add to the single coefficient
\(N_f\), while the vector-multiplet Jacobian produces the denominator in the
chosen canonical coordinate.

\section{Gauge-Invariant Chiral Coordinates}

The meson matrix is the chiral operator
\[
  M^i{}_j=\widetilde Q_j Q^i .
\]
For \(N_f\geq N_c\) there are baryons
\[
  B^{i_1\cdots i_{N_c}}
  =
  \epsilon_{a_1\cdots a_{N_c}}
  Q^{a_1 i_1}\cdots Q^{a_{N_c}i_{N_c}},
\]
and similarly antibaryons \(\widetilde B\).  The classical chiral coordinate
ring is the quotient of the polynomial ring in \(M,B,\widetilde B\) by the
relations obtained from the definitions above.  Quantum dynamics changes this
ring only through Wilsonian chiral coordinates compatible with holomorphy,
global symmetries, and anomalies.

\section{Anomaly-Free R-Symmetry}

Assign
\[
  R(Q)=R(\widetilde Q)=1-\frac{N_c}{N_f},
  \qquad
  R(W_\alpha)=1 .
\]
The \(SU(N_c)^2U(1)_R\) anomaly coefficient is
\[
  N_c+\frac12N_f(R(Q)-1)+\frac12N_f(R(\widetilde Q)-1)=0
\]
in the fundamental normalization where each fundamental contributes
\(\frac12\).  This \(R\)-symmetry is part of the chiral-coordinate selection
data when \(N_f>0\).  If a mass term
\[
  W_{\mathrm{tree}}=m_i{}^jM^i{}_j
\]
is included, then \(m_i{}^j\) is treated as a background chiral coordinate
with the charge needed to make the superpotential have \(R\)-charge \(2\).

\section{Gaugino Condensation}
\label{sec:pure-sym-gaugino-condensation}

For pure \(SU(N_c)\) supersymmetric Yang--Mills, \(b_0=3N_c\).  Let
\[
  S=-\frac{1}{32\pi^2}\operatorname{tr}(W^\alpha W_\alpha)
\]
be the chiral glueball coordinate in the same trace convention.  Its lowest
component is proportional to the gaugino bilinear \(\operatorname{tr}\lambda
\lambda\), so it has engineering dimension \(3\) and gaugino-phase charge
\(2\).

A useful way to keep the nonperturbative status precise is to separate three
statements: the anomaly fixes the possible discrete phases, the Witten index
counts supersymmetric vacua under finite-volume hypotheses, and a glueball
Wilsonian description produces the chiral condensate equation.  A constructive
mass-gap theorem would additionally require a regulator, a Hilbert-space
limit, and spectral estimates for the gauge-invariant local algebra.

\begin{conjecture}[Central Conjecture (Pure \(\mathcal N=1\) supersymmetric Yang--Mills)]
\label{conj:pure-n1-sym-central}
Fix \(N_c\ge2\), gauge group \(SU(N_c)\), theta angle
\(\theta\in\mathbb R/2\pi\mathbb Z\), periodic spin structure in finite-volume
regulators, and the simply connected line-operator lattice with adjoint
dynamical matter.  Pure \(\mathcal N=1\) \(SU(N_c)\) Yang--Mills is conjectured
to have the following continuum infrared structure.
\begin{enumerate}[label=(\roman*)]
\item A four-dimensional supersymmetric continuum gauge theory exists as a
  regulator limit, with a local gauge-invariant superfield algebra, conserved
  supercharges, the expected anomalous \(U(1)_R\) remnant
  \(\mathbb Z_{2N_c}\), and renormalized center-charged line sectors.
\item The infinite-volume theory has \(N_c\) pure supersymmetric phases
  \(\omega_k\), \(k\in\mathbb Z/N_c\mathbb Z\), permuted by
  \(\theta\mapsto\theta+2\pi\) and by the quotient
  \(\mathbb Z_{2N_c}/\mathbb Z_2\).  Each selected phase satisfies cluster
  decomposition.
\item In every pure phase the physical gauge-invariant Hilbert space has a
  strictly positive mass gap, no additional massless chiral coordinates, and
  unbroken supersymmetry.
\item The theory is confining for nonzero \(N_c\)-ality external electric probes
  in the line-sector sense of
  Chapter~\ref{ch:global-confinement-screening-oblique}: the electric one-form
  center symmetry is unbroken and the corresponding renormalized string
  tensions are positive.
\item The discrete chiral symmetry is spontaneously broken as
  \(\mathbb Z_{2N_c}\to\mathbb Z_2\), with a nonzero gaugino condensate
  \begin{equation}
    \langle S\rangle_k
    =
    C\,\Lambda_h^3\exp\!\left(\frac{2\pi\ii k}{N_c}\right),
    \qquad k\in\mathbb Z/N_c\mathbb Z,
    \label{eq:pure-sym-central-condensate-orbit}
  \end{equation}
  for a scheme-dependent nonzero constant \(C\).
\item Domain-wall sectors, BPS wall tensions, confining strings, and massive
  glueball/gluinoball supermultiplets exist as spectral data only after the
  corresponding wall, line, or local-source Hilbert-space sector has been
  constructed.
\end{enumerate}
The gap, confinement, phase purity, cluster decomposition, and nonzero
condensate are separate dynamical claims.  The Witten index, discrete anomaly,
semiclassical small-circle calculation, holomorphic source identity,
Veneziano--Yankielowicz coordinate analysis, and instanton selection rules are
evidence for this package, not individual proofs of it.
\end{conjecture}

\paragraph{Pure \(\mathcal N=1\) SYM evidence ledger.}
The discrete anomaly proves the possible \(\mathbb Z_{2N_c}\) phase orbit but
does not produce a condensate.  The finite-volume Witten index counts protected
zero-energy states in a specified Hilbert-space sector but does not by itself
select cluster vacua, prove a gap, or give a nonzero one-point function.  The
small-circle affine-Toda calculation supplies a controlled semiclassical
\(N_c\)-critical-point model in a compactified chamber, with a spectral
continuation hypothesis needed before it speaks about the four-dimensional
infinite-volume theory.  The VY/source-normalization calculation gives the
allowed chiral root equation once a massive glueball \(F\)-term coordinate is
assumed.  The zero-mode-saturated instanton correlator can test the
\(N_c\)-point chiral selection rule only after regulator, determinant,
instanton-size, collision, and clustering assumptions have been supplied.  All
of these entries support Conjecture~\ref{conj:pure-n1-sym-central}; none
replaces its continuum spectral and phase-purity content.

\paragraph{Pure SYM discrete chiral anomaly.}
In pure \(SU(N_c)\) supersymmetric Yang--Mills, the classical gaugino phase
rotation
\[
  \lambda_\alpha\longmapsto \ee^{\ii\alpha}\lambda_\alpha
\]
is reduced by the adjoint fermion anomaly to
\[
  \mathbb Z_{2N_c}.
\]
If a nonzero condensate \(\langle S\rangle\) forms, this discrete symmetry is
further broken to \(\mathbb Z_2\), giving \(N_c\) possible condensate
phases.
The left-handed adjoint Weyl fermion has anomaly coefficient
\[
  2T(\operatorname{adj})=2N_c
\]
in the instanton normalization used throughout the chapter.  Under the
formal rotation by \(\alpha\), the functional measure shifts the Yang--Mills
angle by
\[
  \theta\longmapsto \theta+2N_c\alpha .
\]
Since \(\theta\) is \(2\pi\)-periodic, the rotation is a genuine symmetry
precisely when
\[
  2N_c\alpha\in2\pi\mathbb Z,
  \qquad\text{that is}\qquad
  \alpha=\frac{\pi\ell}{N_c},\quad \ell\in\mathbb Z .
\]
Modulo \(\alpha\sim\alpha+2\pi\), these rotations form
\(\mathbb Z_{2N_c}\).  The glueball coordinate \(S\) has charge \(2\) under
the same rotation:
\[
  S\longmapsto \ee^{2\ii\alpha}S .
\]
A nonzero value of \(S\) is fixed only when \(\ee^{2\ii\alpha}=1\), which is
the subgroup generated by \(\alpha=\pi\).  Hence the unbroken subgroup is
\(\mathbb Z_2\), and the quotient has \(2N_c/2=N_c\) phases.

\begin{hypothesis}[Pure SYM glueball \(F\)-term description]
\label{hyp:pure-sym-glueball-f-term-description}
The following glueball description is a Wilsonian chiral-sector hypothesis.
Work in a supersymmetric holomorphic scheme for pure \(SU(N_c)\)
Yang--Mills with the glueball coordinate \(S\) normalized as above.  On each
massive chiral branch, assume that the protected \(F\)-term dependence is
locally described by a holomorphic, possibly multivalued, superpotential
\[
  W_{\mathrm{eff}}(S;\Lambda_h^{3N_c})
\]
and that there are no further massless chiral coordinates.  The source
normalization of \(S\) is fixed by
\begin{equation}
  \Lambda_h^{3N_c}
  \frac{\partial W_{\mathrm{eff}}}{\partial \Lambda_h^{3N_c}}
  =
  S .
  \label{eq:pure-sym-glueball-source-normalization}
\end{equation}
Finite holomorphic rescalings of \(S\) or \(\Lambda_h\) change the numerical
constant in the final condensate but not the number of vacua or their
relative phases.
\end{hypothesis}

\paragraph{VY glueball superpotential from the chiral-sector hypothesis.}
Assume Hypothesis~\ref{hyp:pure-sym-glueball-f-term-description}.  Up to the
finite normalization just described, the pure-SYM glueball superpotential may
be written as
\begin{equation}
  W_{\mathrm{VY}}(S)
  =
  S\left(
    \log\frac{\Lambda_h^{3N_c}}{S^{N_c}}+N_c
  \right).
  \label{eq:pure-sym-vy-superpotential}
\end{equation}
It has dimension \(3\), satisfies the source identity
\eqref{eq:pure-sym-glueball-source-normalization}, and its critical points
obey
\begin{equation}
  S^{N_c}=\Lambda_h^{3N_c}.
  \label{eq:pure-sym-vy-critical-equation}
\end{equation}

Set \(L=\Lambda_h^{3N_c}\).  The only dimensionless holomorphic combination
of \(S\) and \(L\) is
\[
  u=\frac{S^{N_c}}{L}.
\]
A dimension-three glueball superpotential therefore has the local form
\[
  W_{\mathrm{eff}}=S f(u)
\]
for a holomorphic function \(f\) on a chosen logarithmic branch.  The source
normalization gives
\[
  S
  =
  L\frac{\partial}{\partial L}\bigl(Sf(u)\bigr)
  =
  -S\,u f'(u).
\]
Thus
\[
  u f'(u)=-1,
\]
so
\[
  f(u)=-\log u+c
\]
on that branch.  Hence
\[
  W_{\mathrm{eff}}
  =
  S\left(\log\frac{L}{S^{N_c}}+c\right).
\]
The constant \(c\) is shifted by a finite holomorphic rescaling of \(L\) or
of \(S\).  The standard Veneziano--Yankielowicz representative is the
normalization \(c=N_c\), giving
\eqref{eq:pure-sym-vy-superpotential}.  Differentiating this representative
with respect to \(S\) yields
\[
  \frac{\partial W_{\mathrm{VY}}}{\partial S}
  =
  \log\frac{\Lambda_h^{3N_c}}{S^{N_c}},
\]
so the \(F_S=0\) equation is
\eqref{eq:pure-sym-vy-critical-equation}.  Finally, both \(S\) and
\(\Lambda_h^3\) have dimension \(3\), and the logarithm is dimensionless;
therefore \(W_{\mathrm{VY}}\) has dimension \(3\).

Choosing a branch of \(\Lambda_h^3\), the solutions of
\eqref{eq:pure-sym-vy-critical-equation} are
\[
  \langle S\rangle_k
  =
  C\,\Lambda_h^3\,
  \exp\!\left(\frac{2\pi\ii k}{N_c}\right),
  \qquad
  k=0,\ldots,N_c-1 .
\]
The constant \(C\) depends on the normalization of \(S\) and on the definition
of \(\Lambda_h\).  In the standard representative of
\eqref{eq:pure-sym-vy-superpotential}, \(C=1\).  At a critical point,
\[
  \frac{\partial^2 W_{\mathrm{VY}}}{\partial S^2}
  =
  -\frac{N_c}{S},
\]
so the glueball critical points are isolated whenever \(S\neq0\).  The
statement with invariant content is the existence of a nonzero chiral
condensate with these relative phases.  Its realization in isolated gapped
pure phases is the corresponding condensate and phase-purity part of
Conjecture~\ref{conj:pure-n1-sym-central}; the finite-dimensional
critical-point calculation records the chiral-coordinate branch equation.

\paragraph{Condensate branches and source normalization.}
Assume Hypothesis~\ref{hyp:pure-sym-glueball-f-term-description}, and assume
that the isolated critical points of
\eqref{eq:pure-sym-vy-critical-equation} are realized as massive chiral
branches of the pure \(SU(N_c)\) theory.  With
\[
  L=\Lambda_h^{3N_c}.
\]
the \(k\)-th branch has
\[
  W_k(L)=N_c\,C\,L^{1/N_c}\ee^{2\pi\ii k/N_c},
  \qquad
  \langle S\rangle_k
  =
  C\,L^{1/N_c}\ee^{2\pi\ii k/N_c},
  \qquad k\in\mathbb Z/N_c\mathbb Z .
\]
The source identity gives
\[
  L\frac{\partial W_k}{\partial L}=\langle S\rangle_k .
\]
Analytic continuation of \(L\) once around the origin sends
\(k\mapsto k+1\).  Thus the \(N_c\) branches form one orbit under the
anomalous chiral symmetry and under a \(2\pi\) theta-angle loop.
The calculation is the following.  At a VY critical point the logarithm in
\eqref{eq:pure-sym-vy-superpotential} vanishes, so the critical value of the
standard representative is \(W_k=N_cS_k\).  The finite normalization
constant \(C\) records the allowed rescaling of \(S\) and of the chosen
branch of \(\Lambda_h^3\).  Differentiating
\[
  W_k(L)=N_c\,C\,L^{1/N_c}\ee^{2\pi\ii k/N_c}
\]
gives
\[
  L\frac{\partial W_k}{\partial L}
  =
  C\,L^{1/N_c}\ee^{2\pi\ii k/N_c}
  =
  \langle S\rangle_k .
\]
The remaining statement is the elementary monodromy of the \(N_c\)-th root:
if \(L\) is carried around a positively oriented loop about the origin, then
\(L^{1/N_c}\) is multiplied by \(\ee^{2\pi\ii/N_c}\).  This shifts the branch
label by one.  The same shift is produced by the generator
\(\alpha=\pi/N_c\) of the anomaly-preserving \(\mathbb Z_{2N_c}\), since
\(S\) has gaugino-phase charge \(2\).

\section{Protected and Spectral Data in Pure SYM}
\label{sec:pure-sym-protected-spectral-data}

The glueball coordinate \(S\) records protected chiral data.  The massive
particle spectrum of pure supersymmetric Yang--Mills is a different object:
it is extracted from the translation representation on the Hilbert space of a
chosen vacuum, or from an equivalent Euclidean transfer matrix after
reconstruction.  This section fixes the precise spectral problem that connects the
chiral branch discussion above to the pure-Yang--Mills spectral discussion of
Section~\ref{sec:lattice-ym-glueballs-flux-observables}.

\begin{definition}[Vacuum spectral problem of pure SYM]
\label{def:pure-sym-vacuum-spectral-problem}
Fix \(N_c\ge2\) and a trace convention for the gauge algebra as in
Section~\ref{sec:bpst-instanton-thooft-vertex}.  A vacuum
spectral problem for pure \(\mathcal N=1\) \(SU(N_c)\) Yang--Mills consists of:
\begin{enumerate}[label=(\roman*)]
\item a regulator or continuum construction that gives a local
  gauge-invariant \(*\)-algebra \(\mathfrak A_{\mathrm{loc}}\) generated by
  smeared components of gauge-invariant superfields;
\item a supersymmetric vacuum state \(\omega_k\) with GNS triple
  \((\mathcal H_k,\Omega_k,\pi_k)\), where \(k\in\mathbb Z/N_c\mathbb Z\)
  labels the chiral branch in the VY/source-normalization discussion above;
\item a strongly continuous translation representation
  \(U_k(a)=\exp(\ii a^\mu P^{(k)}_\mu)\) with joint spectral measure
  \(E_k(\cdot)\), vacuum vector \(P^{(k)}_\mu\Omega_k=0\), and mass operator
  \(M_k^2=-\eta^{\mu\nu}P^{(k)}_\mu P^{(k)}_\nu\);
\item a finite or countable source family
  \(\mathcal O_i\in\mathfrak A_{\mathrm{loc}}\) of gauge-invariant smeared
  component fields, such as the scalar, pseudoscalar, and spinor components
  of \(S\) and their descendants after contact terms have been fixed;
\item the connected spectral measures
  \[
    \rho^{(k)}_{ij}(B)
    =
    \bigl\langle
      \pi_k(\mathcal O_i)\Omega_k,\,
      E_k(B)\,\pi_k(\mathcal O_j)\Omega_k
    \bigr\rangle_{\mathcal H_k}
  \]
  for Borel sets \(B\subset\{p\in\mathbb R^{1,3}:p^0\ge0\}\), with the
  vacuum projection removed when \(\mathcal O_i\) has a nonzero vacuum
  expectation value.
\end{enumerate}
An isolated massive supermultiplet of mass \(m>0\) is represented by an
atomic component of these measures on the mass shell \(p^2=-m^2\).  The
matrix of residues on that atom gives the overlap of the chosen sources with
the one-particle states.
\end{definition}

\paragraph{Chiral \(F\)-term data and quadratic masses.}
Assume that a neighborhood of a branch \(S=S_k\) admits a single-coordinate
two-derivative chiral effective action
\[
  \int\dd^4\theta\,K(S,\overline S)
  +
  \left(\int\dd^2\theta\,W(S)+\mathrm{c.c.}\right),
  \qquad
  g_k=\partial_S\partial_{\overline S}K(S_k,\overline S_k)>0,
\]
and that \(W'(S_k)=0\).  At quadratic order in the fluctuation
\(\sigma=S-S_k\), the complex scalar and its Majorana fermion partner have
common mass
\begin{equation}
  m_k=\frac{|W''(S_k)|}{g_k}.
  \label{eq:pure-sym-single-chiral-mass}
\end{equation}
For the VY representative
\eqref{eq:pure-sym-vy-superpotential},
\begin{equation}
  W''(S_k)=-\frac{N_c}{S_k}.
  \label{eq:pure-sym-vy-second-derivative}
\end{equation}
Thus the protected \(F\)-term fixes the branch values and the holomorphic
second derivative, while the numerical particle masses also require the
Kähler metric and, in the full theory, the complete source-residue matrix of
Definition~\ref{def:pure-sym-vacuum-spectral-problem}.
At a supersymmetric critical point \(W'(S_k)=0\), the scalar potential is
\[
  V(S,\overline S)=
  K^{S\overline S}(S,\overline S)\,|W'(S)|^2 .
\]
Expanding at \(S_k\) gives
\[
  W'(S_k+\sigma)=W''(S_k)\sigma+O(\sigma^2),
  \qquad
  K^{S\overline S}(S_k,\overline S_k)=g_k^{-1}.
\]
The quadratic scalar Lagrangian is therefore
\[
  g_k\,\partial_\mu\sigma\,\partial^\mu\overline\sigma
  -
  g_k^{-1}|W''(S_k)|^2|\sigma|^2 .
\]
After the canonical rescaling \(\sigma_c=g_k^{1/2}\sigma\), the mass-squared
coefficient is \(|W''(S_k)|^2/g_k^2\).  The fermion quadratic term is
\[
  -\frac12 W''(S_k)\psi_S\psi_S+\mathrm{c.c.}
\]
together with the kinetic coefficient \(g_k\); the same canonical rescaling
gives the fermion mass \(|W''(S_k)|/g_k\).  Differentiating
\eqref{eq:pure-sym-vy-superpotential} twice gives
\eqref{eq:pure-sym-vy-second-derivative}.  In the microscopic theory, several
gauge-invariant sources with the same spin, parity, charge conjugation, and
fermion number can have nonzero overlap with the same atom of the spectral
measure, so the physical mass extraction uses the residue matrix rather than
a single holomorphic coordinate alone.

\begin{definition}[Pure-SYM wall data]
\label{def:pure-sym-wall-data}
For two massive supersymmetric branches \(i,j\in\mathbb Z/N_c\mathbb Z\), a
wall data consist of a Hilbert-space or Euclidean functional-integral
construction on \(\mathbb R^{1,2}\times\mathbb R\) with boundary conditions
approaching the \(i\)-th vacuum as \(x^3\to-\infty\) and the \(j\)-th vacuum
as \(x^3\to+\infty\), together with a conserved energy per unit
\(\mathbb R^{1,2}\)-volume.  In a two-derivative chiral effective
description, the wall is represented by a path \(S(x^3)\) in the branch
coordinate space.
\end{definition}

\paragraph{BPS wall tension from branch superpotentials.}
Assume the wall data of Definition~\ref{def:pure-sym-wall-data} are governed
by a two-derivative \(\mathcal N=1\) chiral effective action with positive
Kähler metric \(g_{S\overline S}\), and assume that a wall saturates the
\(\mathcal N=1\) central-charge bound.  Then its tension is
\begin{equation}
  T_{ij}=2\,|W_j-W_i|,
  \label{eq:n1-bps-wall-tension-general}
\end{equation}
where \(W_i\) and \(W_j\) are the branch values of the superpotential.  For
the pure-SYM VY branch values
\[
  W_k=N_c\,C\,\Lambda_h^3\,\ee^{2\pi\ii k/N_c},
\]
the wall between branches \(k\) and \(k+\ell\) has
\begin{equation}
  T_{k,k+\ell}
  =
  4N_c\,|C|\,|\Lambda_h^3|\,
  \left|\sin\frac{\pi\ell}{N_c}\right|.
  \label{eq:pure-sym-wall-sine-tension}
\end{equation}

For a static configuration depending only on \(x^3\), the energy density in
the effective chiral model is
\[
  \mathcal E
  =
  g_{S\overline S}\,\partial_3S\,\partial_3\overline S
  +
  g^{S\overline S}W_S\overline W_{\overline S}.
\]
Let \(\alpha\in\mathbb R\).  Completing the square gives
\[
  \mathcal E
  =
  g_{S\overline S}
  \left|
    \partial_3S-\ee^{\ii\alpha}g^{S\overline S}\overline W_{\overline S}
  \right|^2
  +
  2\,\operatorname{Re}\!\left(
    \ee^{-\ii\alpha}\partial_3 W
  \right).
\]
After integrating over \(x^3\), the last term is
\[
  2\,\operatorname{Re}\!\left(
    \ee^{-\ii\alpha}(W_j-W_i)
  \right).
\]
Choosing \(\alpha=\arg(W_j-W_i)\) gives the bound
\[
  T\ge 2|W_j-W_i|.
\]
Saturation is equivalent to the first-order wall equation
\[
  \partial_3S=\ee^{\ii\alpha}g^{S\overline S}\overline W_{\overline S}.
\]
Substituting the branch values gives
\[
  |W_{k+\ell}-W_k|
  =
  N_c|C|\,|\Lambda_h^3|\,
  \left|\ee^{2\pi\ii(k+\ell)/N_c}-\ee^{2\pi\ii k/N_c}\right|
  =
  2N_c|C|\,|\Lambda_h^3|\,
  \left|\sin\frac{\pi\ell}{N_c}\right|,
\]
which proves \eqref{eq:pure-sym-wall-sine-tension}.

Thus the sine formula is a consequence of the effective chiral wall chart and
the saturation hypothesis.  The nontrivial QFT problem is the construction of
the wall sector and the justification that this two-derivative description
with the stated branch superpotential controls its central charge.  The
square-completion records only the effective-chart consequence of those data.

Pure-SYM confining string tensions are additional line-operator spectral data.
They are defined by the area-law exponent of center-charged Wilson-loop or
't Hooft-loop correlators, or equivalently by a transfer-matrix flux-sector
energy in a regulator that realizes the center sectors.  The wall formula
\eqref{eq:pure-sym-wall-sine-tension} is protected by the chiral
superpotential under the stated effective-action hypotheses; the flux-string
tension requires the nonchiral dynamics encoded in the spectral data.

\paragraph{Pure-SYM one-instanton zero-mode test.}
\label{par:pure-sym-one-instanton-zero-mode-test}
Work in the charge-one \(SU(N_c)\) instanton sector in the BPST
normalization of Section~\ref{sec:bpst-instanton-thooft-vertex}.  Assume a
semiclassical collective-coordinate separation in which the exact adjoint
fermion zero modes are represented by Grassmann variables and must be
saturated by separated operator insertions.  Then
\begin{equation}
  \left\langle S(x_1)\cdots S(x_m)\right\rangle_{k=1}^{\mathrm{sep}}=0,
  \qquad m<N_c,
  \label{eq:pure-sym-instanton-too-few-s-insertions}
\end{equation}
before contact terms or boundary contributions are considered.  The first
separated chiral correlator whose fermion degree can saturate the
charge-one instanton measure is the \(N_c\)-point correlator
\begin{equation}
  \left\langle S(x_1)\cdots S(x_{N_c})\right\rangle_{k=1}^{\mathrm{sep}}.
  \label{eq:pure-sym-instanton-first-saturated-correlator}
\end{equation}
Its protected charge and dimension assignment is the same assignment as the VY
equation \(S^{N_c}\sim \Lambda_h^{3N_c}\).

The Dirac index in the adjoint representation gives
\[
  n_\lambda=2T(\operatorname{adj})\,k=2N_c
\]
left-handed gaugino zero modes for a charge-one instanton.  Let
\(\eta_1,\ldots,\eta_{2N_c}\) denote their Grassmann coefficients.  The
zero-mode part of the collective-coordinate measure contains
\[
  d\eta_{2N_c}\cdots d\eta_1 .
\]
The lowest component of \(S\) is quadratic in the gaugino field.  Hence a
separated insertion of \(S\) is a polynomial of Grassmann degree at most
two in the \(\eta_a\)'s, and a product of \(m\) such insertions has degree at
most \(2m\).  A Berezin integral over \(2N_c\) variables extracts only the
coefficient of \(\eta_1\cdots\eta_{2N_c}\).  If \(m<N_c\), that coefficient
is absent, proving \eqref{eq:pure-sym-instanton-too-few-s-insertions}.

For \(m=N_c\), the Grassmann degree can be \(2N_c\).  This is a necessary
zero-mode condition, not yet a proof of a nonzero coefficient; the
spacetime dependence, nonzero-mode determinant, instanton-size integration,
and possible boundary terms still have to be controlled.  The holomorphic
scale factor of a charge-one pure-SYM instanton is
\(\Lambda_h^{3N_c}\), while \(S^{N_c}\) has engineering dimension \(3N_c\)
and gaugino-phase charge \(2N_c\).  Thus the first zero-mode-saturated
chiral correlator has precisely the dimension and anomalous-charge assignment
of the VY critical equation.

\paragraph{Saturated pure-SYM Berezin coefficient.}
The following exterior-algebra calculation is not a separate QFT theorem; it
records the sign and factorial convention needed by the instanton
zero-mode test.  Keep the assumptions and notation of
the one-instanton zero-mode test above.  At fixed
bosonic instanton collective coordinate \(\mathfrak m\), let
\[
  \eta_1,\ldots,\eta_{2N_c}
\]
be an oriented basis of the adjoint gaugino zero-mode Grassmann variables,
with Berezin measure
\[
  \dd\eta_{2N_c}\cdots\dd\eta_1 .
\]
For a separated insertion point \(x_a\), write the zero-mode part of the
lowest component of \(S(x_a)\) as the quadratic form
\begin{equation}
  S(x_a)\big|_{\mathrm{zm}}
  =
  \sum_{1\leq p<q\leq 2N_c}
  K_a^{pq}(x_a;\mathfrak m)\,\eta_p\eta_q ,
  \label{eq:pure-sym-s-insertion-zero-mode-quadratic-form}
\end{equation}
where the coefficients \(K_a^{pq}\) are ordinary complex numbers after the
bosonic collective coordinate and the zero-mode basis have been fixed.  Define
the saturated Berezin coefficient
\begin{equation}
  \mathfrak B_{N_c}(K_1,\ldots,K_{N_c})
  =
  [\eta_1\cdots\eta_{2N_c}]
  \prod_{a=1}^{N_c}
  \left(
    \sum_{p<q}K_a^{pq}\eta_p\eta_q
  \right),
  \label{eq:pure-sym-saturated-berezin-coefficient}
\end{equation}
where \([\eta_1\cdots\eta_{2N_c}]\) denotes coefficient extraction in the
ordered monomial \(\eta_1\cdots\eta_{2N_c}\).  Equivalently,
\begin{equation}
  \mathfrak B_{N_c}
  =
  \sum_{\substack{p_a<q_a\\
        \bigsqcup_{a=1}^{N_c}\{p_a,q_a\}
        =\{1,\ldots,2N_c\}}}
  \operatorname{sgn}(p_1,q_1,\ldots,p_{N_c},q_{N_c})
  \prod_{a=1}^{N_c}K_a^{p_aq_a}.
  \label{eq:pure-sym-saturated-berezin-expansion}
\end{equation}
The fixed-\(\mathfrak m\) zero-mode integral of the \(N_c\)-point insertion is
\begin{equation}
  \int \dd\eta_{2N_c}\cdots\dd\eta_1
  \prod_{a=1}^{N_c}S(x_a)\big|_{\mathrm{zm}}
  =
  \mathfrak B_{N_c}(K_1,\ldots,K_{N_c}) .
  \label{eq:pure-sym-fixed-moduli-berezin-integral}
\end{equation}
Thus the saturated instanton calculation requires a nonzero top exterior
component of the \(N_c\) two-forms \(K_a\).  In the canonical test case
\[
  K_a=\eta_{2a-1}\eta_{2a},
  \qquad a=1,\ldots,N_c,
\]
the coefficient is \(1\).  If instead every insertion has the same canonical
two-form
\[
  \Omega_0=\sum_{a=1}^{N_c}\eta_{2a-1}\eta_{2a},
\]
then the coefficient of \(\Omega_0^{N_c}\) is \(N_c!\).  These factorial and
sign conventions are part of the instanton calculation; they are not fixed
by dimension, holomorphy, or the index.

Indeed, Berezin integration over the zero-mode variables extracts the coefficient of
the oriented top monomial:
\[
  \int \dd\eta_{2N_c}\cdots\dd\eta_1\,
  \eta_1\cdots\eta_{2N_c}=1 .
\]
Substituting
\eqref{eq:pure-sym-s-insertion-zero-mode-quadratic-form} into the product of
insertions therefore gives
\eqref{eq:pure-sym-fixed-moduli-berezin-integral} directly.  Expanding the
product, a monomial contributes only if every variable
\(\eta_1,\ldots,\eta_{2N_c}\) occurs exactly once.  Ordering that monomial to
\(\eta_1\cdots\eta_{2N_c}\) produces the sign in
\eqref{eq:pure-sym-saturated-berezin-expansion}.  If a variable is missing or
repeated, the coefficient of the top monomial is zero.

For \(K_a=\eta_{2a-1}\eta_{2a}\), the product is already the ordered top
monomial, hence the coefficient is \(1\).  For
\(\Omega_0=\sum_a\eta_{2a-1}\eta_{2a}\), the only nonzero terms in
\(\Omega_0^{N_c}\) choose each pair exactly once; the \(N_c!\) permutations
of these commuting degree-two pair factors all contribute with positive sign.
This proves the displayed coefficient and records the convention-sensitive
normalization that later gets absorbed into the nonzero constant
\(\kappa_{\mathrm{inst}}\).

\paragraph{Topological-sector selection for \(S\)-correlators.}
Fix an instanton number \(\nu\in\mathbb Z_{>0}\).  Work in a self-dual
charge-\(\nu\) sector of pure \(SU(N_c)\) SYM, with the same BPST/index
normalization as above.  Assume that the exact adjoint gaugino zero modes are
separated as Grassmann variables and that, in the separated chiral correlator
under consideration, the only insertions carrying these variables are the
displayed \(S\)-insertions.  Then
\begin{equation}
  \left\langle S(x_1)\cdots S(x_m)\right\rangle_{\nu}^{\mathrm{sep}}=0,
  \qquad m<\nu N_c,
  \label{eq:pure-sym-topological-sector-too-few-s}
\end{equation}
at the zero-mode Berezin level.  The first possible saturated chiral
correlator in the charge-\(\nu\) sector has
\[
  m=\nu N_c
\]
insertions and carries the same dimension and anomalous gaugino-phase charge
as \(\Lambda_h^{3\nu N_c}\).  Consequently the \(N_c\)-point correlator used
in Hypothesis~\ref{hyp:pure-sym-zero-mode-saturated-instanton-correlator} is
a charge-one test: sectors with \(\nu\geq2\) leave unsaturated adjoint zero
modes in that channel.  Anti-self-dual sectors have the conjugate chirality
zero modes and are paired with anti-chiral \(\overline S\)-insertions rather
than with the holomorphic \(S\)-correlator above.

The selection rule is zero-mode degree counting.
The adjoint Dirac index gives
\[
  n_\lambda(\nu)=2T(\operatorname{adj})\,\nu=2N_c\nu
\]
left-handed gaugino zero modes in a self-dual charge-\(\nu\) sector.  A
separated insertion of \(S\) is quadratic in the left-handed gaugino
zero-mode variables, so \(m\) such insertions have Grassmann degree at most
\(2m\).  The zero-mode Berezin integral in the charge-\(\nu\) sector extracts
the coefficient of a monomial of degree \(2N_c\nu\).  If \(m<\nu N_c\), that
top-degree coefficient is absent, proving
\eqref{eq:pure-sym-topological-sector-too-few-s}.

For \(m=\nu N_c\), the fermion degree can match the top degree.  This is only
the algebraic saturation condition; the corresponding coefficient and the
bosonic moduli integral remain separate analytic inputs, as in
Hypothesis~\ref{hyp:pure-sym-zero-mode-saturated-instanton-correlator}.  The
dimension of \(S^{\nu N_c}\) is \(3\nu N_c\), and its gaugino-phase charge is
\(2\nu N_c\), matching the holomorphic scale factor
\(\Lambda_h^{3\nu N_c}\) and the index count of the zero modes.  Setting
\(m=N_c\) forces \(\nu=1\) for equality.  For \(\nu\geq2\), the same
\(N_c\)-point insertion has degree \(2N_c\), leaving
\(2(\nu-1)N_c\) adjoint zero modes unsaturated.

The chirality statement is the conjugate form of the same index theorem:
anti-self-dual sectors carry the opposite chiral zero modes.  The holomorphic
operator \(S\) is built from the chiral field strength \(W_\alpha W^\alpha\),
whereas the conjugate zero modes are supplied by \(\overline S\).  Thus they
belong to the anti-chiral conjugate calculation, apart from contact terms,
source insertions, or boundary effects outside the separated chiral
selection rule.

\begin{hypothesis}[Zero-mode-saturated pure-SYM instanton correlator]
\label{hyp:pure-sym-zero-mode-saturated-instanton-correlator}
The following is the additional analytic input required if one wants to use
the original instanton logic as evidence for the pure-SYM condensate.  In a
supersymmetric holomorphic regulator, assume that the separated
\(N_c\)-point chiral correlator
\eqref{eq:pure-sym-instanton-first-saturated-correlator} receives a
well-defined charge-one instanton contribution
\begin{equation}
  \left\langle S(x_1)\cdots S(x_{N_c})\right\rangle_{k=1}^{\mathrm{sep}}
  =
  \kappa_{\mathrm{inst}}\Lambda_h^{3N_c},
  \qquad \kappa_{\mathrm{inst}}\neq0 ,
  \label{eq:pure-sym-zero-mode-saturated-correlator-input}
\end{equation}
for separated points in the chiral sector, in the same finite normalization
of \(S\) and \(\Lambda_h\) used above.  This hypothesis includes control of
the saturated Berezin coefficient
\eqref{eq:pure-sym-saturated-berezin-coefficient} after integration over
bosonic collective coordinates, the instanton-size integral, the nonzero-mode
determinant, collision singularities, and boundary strata.  In un-Higgsed
pure SYM these analytic controls are not supplied by the elementary zero-mode
count alone.
\end{hypothesis}

\begin{remark}[Cluster extraction is a separate assumption]
\label{rem:pure-sym-instanton-cluster-extraction}
Assume Hypothesis~\ref{hyp:pure-sym-zero-mode-saturated-instanton-correlator}
and assume the infinite-volume theory has a massive pure phase
\(\omega_\ell\) with cluster decomposition and one-point value
\[
  \omega_\ell(S)=s_\ell .
\]
If the separated chiral \(N_c\)-point function is identified with the
large-separation limit in that pure phase, then
\begin{equation}
  s_\ell^{N_c}
  =
  \kappa_{\mathrm{inst}}\Lambda_h^{3N_c}.
  \label{eq:pure-sym-cluster-root-equation}
\end{equation}
Consequently the instanton \(N_c\)-point calculation can at most supply the
root equation for the condensate.  The choice of a particular root and the
existence of the massive pure phases require the anomaly, index, and
massive-branch inputs stated in this section.

Choose separated points \(x_1,\ldots,x_{N_c}\), and translate them so that
their mutual distances go to infinity.  Cluster decomposition in the pure
phase gives
\[
  \lim_{\min_{a\neq b}|x_a-x_b|\to\infty}
  \omega_\ell\!\left(S(x_1)\cdots S(x_{N_c})\right)
  =
  \prod_{a=1}^{N_c}\omega_\ell(S)
  =
  s_\ell^{N_c}.
\]
If this large-separation limit is the protected chiral correlator
\eqref{eq:pure-sym-zero-mode-saturated-correlator-input}, then
\eqref{eq:pure-sym-cluster-root-equation} follows.  The equation determines
only the \(N_c\)-th power.  Its \(N_c\) roots are then organized by the
\(\mathbb Z_{2N_c}\) anomaly and the \(\theta\)-monodromy described in
the VY/source-normalization discussion above; no
single-instanton Berezin integral selects one branch by itself.
\end{remark}

The component condensate \(\langle\operatorname{tr}\lambda\lambda\rangle_k\)
is obtained from \(\langle S\rangle_k\) by the fixed component convention in
the definition of \(S\).  The phase information above is convention
independent.  In a massive infinite-volume vacuum, cluster decomposition
would give
\[
  \lim_{|x|\to\infty}
  \langle S(x)S(0)\rangle_k
  =
  \langle S\rangle_k^2 .
\]
This is an infinite-volume statement about a chosen pure phase.  A
finite-volume ground-state basis can instead diagonalize the discrete
symmetry, and the passage to cluster vacua is part of the massive-branch
hypothesis rather than a consequence of the index alone.

\section{Witten Index Of Pure Supersymmetric Yang--Mills}

\begin{hypothesis}[Finite-volume pure-SYM index problem]
\label{hyp:pure-sym-finite-volume-index-problem}
Fix \(N_c\geq2\) and take the gauge group to be the simply connected group
\(SU(N_c)\).  Put the theory on a spatial three-torus \(T^3_L\) with
periodic spin structure for the gaugino, and impose Gauss' law in the
trivial discrete-flux sector.  The index is computed by a regulated
Hamiltonian system whose supersymmetry, gauge invariance, and fermion parity
are exact at finite regulator.  The physical Hilbert space
\(\mathcal H_0(T^3_L)\) has a self-adjoint Hamiltonian \(H\), closed
supercharges \(Q,Q^\dagger\), and a fermion-parity operator \((-1)^F\) such
that
\[
  H=\frac12\{Q,Q^\dagger\},
  \qquad
  \{(-1)^F,Q\}=\{(-1)^F,Q^\dagger\}=0 .
\]
The heat operator \(\ee^{-\beta H}\) is trace class in the regulated
finite-volume problem, and the zero-energy subspace is finite-dimensional.
Changing the global form of the gauge group or turning on 't Hooft fluxes
changes the Hilbert-space sector and is not part of the displayed
\(SU(N_c)\) index.
\end{hypothesis}

Under Hypothesis~\ref{hyp:pure-sym-finite-volume-index-problem}, the
finite-volume index is
\[
  \mathcal I(\beta,L)
  =
  \operatorname{Tr}_{\mathcal H_0(T^3_L)}
  (-1)^F\ee^{-\beta H}.
\]

\paragraph{Finite-volume graded-trace identity.}
Assume Hypothesis~\ref{hyp:pure-sym-finite-volume-index-problem}.  The
finite-dimensional supersymmetry algebra then gives
\[
  \mathcal I(\beta,L)
  =
  \dim\ker H|_{\mathrm{bosonic}}
  -
  \dim\ker H|_{\mathrm{fermionic}}
\]
and is independent of \(\beta\).  Invariance under a continuous change of
coupling, volume, or regulator follows provided no zero-energy state is lost
to a continuum, no state enters from a continuum, and the deformation
preserves the same supercharge, spin structure, global form, and flux sector.

If \(H|\psi\rangle=E|\psi\rangle\) with \(E>0\), then at least one of
\(Q|\psi\rangle\) and \(Q^\dagger|\psi\rangle\) is nonzero; otherwise
\[
  E\|\psi\|^2
  =
  \frac12\bigl(\|Q|\psi\rangle\|^2+\|Q^\dagger|\psi\rangle\|^2\bigr)
  =
  0 .
\]
The nonzero image has the same energy and opposite fermion parity.  Thus the
positive-energy spectrum contributes zero to the supertrace.  Only
\(\ker H\) remains, giving the displayed formula and \(\beta\)-independence.

For a continuous family of finite-volume Hamiltonians satisfying the same
algebra, positive-energy levels can cross or split only in supersymmetric
boson-fermion multiplets.  The integer supertrace can change only if the
definition of the Hilbert-space sector changes or if zero-energy states
escape to, or arrive from, a continuum.  This is the analytic caveat: the
graded trace identity is algebraic, while deformation invariance of a
continuum quantum field theory is a statement about the spectral behavior of
the regulated family.

\paragraph{Finite-volume symmetry basis versus cluster branches.}
The following finite-dimensional calculation is only an algebraic model for the
zero-energy sector after the massive-branch hypotheses have been imposed.
Let \(V_0\) be an \(N_c\)-dimensional zero-energy space on which all states
have the same fermion parity \(\varepsilon=\pm1\).  Suppose a unitary
operator \(U\), representing the generator of the quotient
\(\mathbb Z_{2N_c}/\mathbb Z_2\), acts on a branch basis
\(\{|k\rangle\}_{k\in\mathbb Z/N_c\mathbb Z}\) by
\[
  U|k\rangle=|k+1\rangle .
\]
Suppose also that the chiral coordinate \(S\) is diagonal in this branch
basis,
\[
  S|k\rangle=s\,\omega^k|k\rangle,
  \qquad
  \omega=\exp(2\pi\ii/N_c),
  \qquad s\neq0 .
\]
Then
\begin{equation}
  \operatorname{Tr}_{V_0}(-1)^F=\varepsilon N_c,
  \qquad
  \operatorname{Tr}_{V_0}\!\left(U^p(-1)^F\right)=0
  \quad (p\not\equiv0\!\!\pmod {N_c}),
  \label{eq:pure-sym-twined-index-regular-representation}
\end{equation}
and the \(U\)-eigenvectors
\[
  |r\rangle
  =
  \frac{1}{\sqrt{N_c}}
  \sum_{k=0}^{N_c-1}\omega^{-rk}|k\rangle,
  \qquad r\in\mathbb Z/N_c\mathbb Z ,
\]
satisfy
\begin{equation}
  U|r\rangle=\omega^r|r\rangle,
  \qquad
  \langle r|S|r\rangle=0,
  \qquad
  S|r\rangle=s\,|r-1\rangle .
  \label{eq:pure-sym-finite-volume-symmetry-basis}
\end{equation}
Thus a finite-volume basis that diagonalizes the exact discrete symmetry
need not display a nonzero one-point condensate.  The nonzero values
\(\langle k|S|k\rangle=s\omega^k\) belong to the cluster branch basis, which
is an infinite-volume pure-phase idealization requiring the additional
massive-branch and cluster hypotheses, beyond the ordinary Witten index.

In the branch basis, \(U^p\) is the permutation \(k\mapsto k+p\).  If
\(p\not\equiv0\pmod{N_c}\), this permutation has no fixed basis vector and
therefore has trace zero.  If \(p\equiv0\), the trace is \(N_c\).  Since
\((-1)^F=\varepsilon\) on \(V_0\), this proves
\eqref{eq:pure-sym-twined-index-regular-representation}.

The displayed Fourier vectors obey
\[
  U|r\rangle
  =
  \frac{1}{\sqrt{N_c}}
  \sum_k \omega^{-rk}|k+1\rangle
  =
  \omega^r|r\rangle .
\]
Their diagonal \(S\)-matrix element is
\[
  \langle r|S|r\rangle
  =
  \frac{s}{N_c}\sum_{k=0}^{N_c-1}\omega^k
  =
  0 ,
\]
because \(\omega\neq1\) and \(\omega^{N_c}=1\).  Finally,
\[
  S|r\rangle
  =
  \frac{s}{\sqrt{N_c}}
  \sum_k \omega^{(1-r)k}|k\rangle
  =
  s\,|r-1\rangle .
\]
The operator \(S\) therefore changes the discrete chiral charge sector
rather than producing a diagonal order parameter in the symmetry-eigenstate
basis.  The branch basis diagonalizes \(S\) but not \(U\), exactly as a
symmetry-breaking basis should.

\begin{hypothesis}[Small-circle affine-Toda input]
\label{hyp:pure-sym-small-circle-affine-toda-input}
The following semiclassical computation is used as a controlled chiral-sector
model for the same \(SU(N_c)\) theory with periodic fermions on
\(\mathbb R^3\times S^1_R\).  Work at small \(R\) in a weak-coupling region
with a generic Wilson line breaking
\[
  SU(N_c)\longrightarrow U(1)^{N_c-1}.
\]
The Wilson-line scalars and dual photons are assembled into periodic chiral
coordinates \(Y_i\), \(i=1,\ldots,N_c\), with
\(\sum_iY_i=0\).  In the basis
\[
  \alpha_i=e_i-e_{i+1},\quad i=1,\ldots,N_c-1,
  \qquad
  \alpha_0=e_{N_c}-e_1,
\]
the \(N_c-1\) fundamental monopole-instantons and the affine monopole each
have exactly two gaugino zero modes and contribute a chiral superpotential
term.  Their product has the topological charge and holomorphic weight of
one four-dimensional instanton, denoted by the compactified scale coordinate
\(\eta\).  Effects with additional fermion zero modes, nonchiral
determinants, and boundary terms are assumed not to change the protected
critical-point count.  After the chiral variables are gapped, the remaining
three-dimensional theory may be placed in finite spatial volume, and the
finite-dimensional local oscillator calculation below applies to each
nondegenerate massive chiral critical point.
\end{hypothesis}

Under Hypothesis~\ref{hyp:pure-sym-small-circle-affine-toda-input}, the
monopole terms generate an affine-Toda superpotential of the form
\[
  W_{\mathrm{aff}}
  =
  \sum_{i=1}^{N_c-1}\exp(Y_i-Y_{i+1})
  +
  \eta\,\exp(Y_{N_c}-Y_1),
\]
where \(\eta\) is the four-dimensional holomorphic scale coordinate in the
compactified normalization.  Define the nonzero variables
\[
  x_i=\exp(Y_i-Y_{i+1}),\quad i=1,\ldots,N_c-1,
  \qquad
  x_{N_c}=\eta\,\exp(Y_{N_c}-Y_1).
\]
The constraint is
\[
  x_1x_2\cdots x_{N_c}=\eta .
\]

\paragraph{Affine-Toda critical-point count.}
Assume Hypothesis~\ref{hyp:pure-sym-small-circle-affine-toda-input} and
\(\eta\neq0\).  The critical equations of \(W_{\mathrm{aff}}\) on the
constraint hypersurface
\[
  x_1\cdots x_{N_c}=\eta
\]
have exactly \(N_c\) isolated solutions:
\[
  x_1=x_2=\cdots=x_{N_c}.
\]
Equivalently, if \(x\) is the common value, then
\[
  x^{N_c}=\eta .
\]

Write variations tangent to the constraint as
\[
  \delta x_i=x_i\,\delta y_i,
  \qquad
  \sum_{i=1}^{N_c}\delta y_i=0 .
\]
The first variation of \(W_{\mathrm{aff}}=\sum_i x_i\) is
\[
  \delta W_{\mathrm{aff}}
  =
  \sum_{i=1}^{N_c}x_i\,\delta y_i .
\]
This vanishes for every vector \((\delta y_i)\) with zero sum if and only if
all \(x_i\) are equal.  The constraint then gives \(x^{N_c}=\eta\), which has
\(N_c\) solutions when \(\eta\neq0\).

The solutions are isolated.  On the tangent hyperplane
\(\sum_i\delta y_i=0\), the quadratic term at a critical point is
\[
  \frac{x}{2}\sum_{i=1}^{N_c}(\delta y_i)^2 .
\]
Since \(x\neq0\), the bilinear form \(\sum_i(\delta y_i)^2\) is
nondegenerate on the hyperplane.  Thus each critical point is Morse in the
complex holomorphic sense.

\begin{proposition}[Local index of a nondegenerate chiral critical point]
\label{prop:pure-sym-local-chiral-critical-index}
Consider a finite-dimensional supersymmetric quantum mechanics obtained by
keeping \(n\) complex chiral zero modes \(z^a\), a positive Hermitian metric
at a critical point \(p\), and a holomorphic superpotential \(W\).  Assume
that
\[
  \partial_a W(p)=0,
  \qquad
  H_{ab}=\partial_a\partial_bW(p)
\]
is an invertible complex symmetric matrix, and that the regulator localizes
the low-energy contribution to a small ball around \(p\) without introducing
extra boundary zero modes.  Then \(p\) contributes
\[
  \mathcal I_{\mathrm{loc}}(p)=+1
\]
to the graded trace.  In particular, the local contribution in the
affine-Toda chamber is not the real Morse sign of \(\operatorname{Re}W\);
the holomorphic chiral complex has a canonical complex orientation.
\end{proposition}

\begin{proof}
Choose normal holomorphic coordinates for the Hermitian metric at \(p\), and
discard higher terms in \(W\).  The discarded terms give a compact
perturbation of the local oscillator as long as no boundary zero mode is
created, so the local index is unchanged by the finite-dimensional
graded-trace identity.  Since \(H\) is complex symmetric and invertible,
Takagi factorization gives a unitary matrix \(U\) and nonzero singular values
\(\mu_a>0\) such that
\[
  H=U^T\operatorname{diag}(\mu_1,\ldots,\mu_n)U .
\]
After this holomorphic change of variables, the quadratic model is a tensor
product of one-dimensional chiral oscillators with
\[
  W_{\mathrm{quad}}
  =
  \frac12\sum_{a=1}^n \mu_a (z^a)^2 .
\]
For one factor, the supersymmetric ground-state equation is the Gaussian
equation
\[
  \left(\frac{\partial}{\partial \bar z}
        +\mu z\right)\Psi_0=0,
  \qquad
  \Psi_0(z,\bar z)=\exp(-\mu |z|^2)
\]
up to normalization and the harmless convention-dependent factor in the
oscillator frequency.  This state is normalizable, unique, and lies in the
even fermion sector.  Every excited oscillator state is paired with a state
of opposite fermion parity by the local supercharge.  Hence one complex
factor has local supertrace \(+1\), and the tensor product of \(n\) factors
also has local supertrace \(+1\).

The statement about signs follows from the same normal form.  As a real
Morse function, \(\operatorname{Re}(\mu z^2/2)\) has one positive and one
negative Hessian direction for each complex coordinate.  That real Morse
sign belongs to the de Rham Morse complex of \(\operatorname{Re}W\), not to
the holomorphic chiral oscillator used in the supersymmetric gauge-theory
index.
\end{proof}

\begin{hypothesis}[Index continuation and massive-branch interpretation]
\label{hyp:pure-sym-index-continuation-massive-branch}
The \(SU(N_c)\) index obtained in the small-circle semiclassical chamber is
identified with the four-dimensional finite-volume index only when no
zero-energy state leaks through a continuum and the spin structure, global
form, and flux sector are unchanged.  The infinite-volume limit is assumed
to have \(N_c\) massive pure phases, related by the anomalous
\(\mathbb Z_{2N_c}\) chiral symmetry and satisfying cluster decomposition.
These are precisely the phase-purity and cluster parts of
Conjecture~\ref{conj:pure-n1-sym-central} when used as four-dimensional
infrared input.  Confinement and a mass gap require the separate spectral
construction described in Section~\ref{sec:pure-sym-protected-spectral-data}.
\end{hypothesis}

\paragraph{Pure \(SU(N_c)\) index and condensate relation.}
Assume Hypotheses~\ref{hyp:pure-sym-finite-volume-index-problem},
\ref{hyp:pure-sym-small-circle-affine-toda-input}, and
\ref{hyp:pure-sym-index-continuation-massive-branch}.  Then
\[
  \mathcal I_{SU(N_c)}=N_c .
\]
If, in addition, Hypothesis~\ref{hyp:pure-sym-glueball-f-term-description}
realizes the massive chiral branches through the glueball coordinate \(S\),
then the \(N_c\) branches have condensates
\[
  \langle S\rangle_k
  =
  C\,\Lambda_h^3\ee^{2\pi\ii k/N_c},
  \qquad k=0,\ldots,N_c-1,
\]
and the discrete chiral symmetry is broken as
\(\mathbb Z_{2N_c}\to\mathbb Z_2\) in each pure phase.

The affine-Toda computation gives \(N_c\) isolated supersymmetric critical
points in the small-circle chamber.  The finite-volume graded-trace identity
turns this critical-point count into the index in that chamber, because the
critical points are isolated and massive in the chiral variables of the
semiclassical approximation, and
Proposition~\ref{prop:pure-sym-local-chiral-critical-index} gives local
contribution \(+1\) at each of them.  The continuation hypothesis then
identifies the same integer with the four-dimensional \(SU(N_c)\)
finite-volume index.

The condensate statement uses an input beyond the index: the additional
glueball \(F\)-term hypothesis.  Under that hypothesis,
the VY/source-normalization branch analysis gives precisely
the displayed \(N_c\)-fold orbit of \(\langle S\rangle\).  Since \(S\) has
gaugino-phase charge \(2\), a nonzero value of \(\langle S\rangle_k\) leaves
only the \(\alpha=\pi\) subgroup unbroken, namely \(\mathbb Z_2\).
This paragraph is an evidence ledger for
Conjecture~\ref{conj:pure-n1-sym-central}: it assembles the protected count,
small-circle critical points, and chiral-source orbit, while the infinite-volume
gap, confinement, and phase-purity claims remain part of the central
conjecture.

For a simple simply connected gauge group \(G\), the expected index is the
dual Coxeter number \(h^\vee_G\).  The monograph does not use that general
statement here; the detailed derivation above is the \(SU(N_c)\) case, where
\(h^\vee_{SU(N_c)}=N_c\).  The delicate point is still the route from the
regulated finite-volume computation to the four-dimensional infinite-volume
theory: the index protects a signed count only under the stated spectral and
boundary-condition hypotheses, while the condensate requires the additional
massive-branch and chiral-source input.

\section{ADS Superpotential}

The ADS superpotential enters here as a Wilsonian chiral quantity obtained
from a controlled semiclassical seed and then continued holomorphically.
Symmetry and holomorphy are powerful only after one knows which chiral
quantity is being computed, on which patch it is controlled, and which
boundary terms have been excluded or absorbed into the holomorphic scale.
The constructive seed is the \(N_f=N_c-1\) theory on the
maximal-rank Higgs patch.  There the gauge group is completely broken, the
instanton size integral is cut off at large size by the Higgs expectation
values, and the charge-one collective-coordinate integral can be reduced to
finite-dimensional bosonic and fermionic zero-mode algebra together with a
regulated determinant over nonzero modes.

The logical order used below is therefore the following.  First, count the
BPST collective coordinates in the normalization fixed earlier and identify
the exact two gaugino modes that can become the \(d^2\theta\) measure.
Second, use the microscopic Yukawa coupling in the Higgs background to lift
all remaining matter and non-Goldstino gaugino zero modes, and compute the
corresponding Berezin determinant instead of replacing it by an uncomputed
``zero-mode saturation'' phrase.  Third, check the real instanton-size
integral that shows how the Higgs vevs control the large-size endpoint.  Fourth, use
holomorphy, special-flavor invariance, and the scaling degree to identify the
remaining chiral function as \(C/\det M\).  Finally, only after this
semiclassical seed is in place, holomorphic mass decoupling propagates the
result to \(N_f<N_c-1\).

This order matters conceptually.  The explicit instanton calculation supplies
the existence of a nonzero \(F\)-term, the scale exponent, the exact
zero-mode selection rule, and the meson-patch denominator in a regime where
the functional integral has a semiclassical meaning.  Holomorphy then extends
that chiral input.  Confinement, a mass gap, and a complete Hilbert-space
description require the Wilsonian and infrared hypotheses separated below.

Assume \(N_f<N_c\) and consider the region where \(M\) has maximal rank.  The
holomorphic superpotential compatible with flavor symmetry, the anomaly-free
\(R\)-symmetry, and the scale coordinate has the form
\[
  W_{\mathrm{ADS}}
  =
  C_{N_c,N_f}
  \left(
  \frac{\Lambda_h^{3N_c-N_f}}{\det M}
  \right)^{\frac{1}{N_c-N_f}} .
\]
Its mass dimension is \(3\), because
\[
  [\Lambda_h^{3N_c-N_f}]=3N_c-N_f,\qquad
  [\det M]=2N_f,
\]
and
\[
  \frac{3N_c-N_f-2N_f}{N_c-N_f}=3 .
\]
Its \(R\)-charge is \(2\), since \(R(M)=2(1-N_c/N_f)\), hence
\[
  R(\det M)=2N_f-2N_c,
  \qquad
  R\!\left((\det M)^{-1/(N_c-N_f)}\right)=2 .
\]
The coefficient is fixed by matching to a semiclassical instanton
calculation for \(N_f=N_c-1\) and by holomorphic decoupling for smaller
\(N_f\), once the instanton measure and scale convention have been specified.

\begin{hypothesis}[SQCD one-instanton calculus input]
\label{hyp:sqcd-one-instanton-calculus}
For the direct instanton derivation at \(N_f=N_c-1\), the following analytic
inputs are assumed.  The Euclidean functional integral is regulated in a
supersymmetric holomorphic scheme preserving
\(SU(N_f)_L\times SU(N_f)_R\times U(1)_B\) and the anomaly-free
\(U(1)_R\).  The vacuum is taken on the maximal-rank Higgs patch, so the
gauge group is completely broken and the instanton size integral is cut off
at large size by the Higgs scale.  The collective-coordinate change of
variables separates exact zero modes from nonzero modes, with the topological
charge, \(\theta\)-angle, and coupling normalization fixed by
Section~\ref{sec:bpst-instanton-thooft-vertex}.  Boundary terms from the
small-size endpoint are absorbed into the holomorphic scale coordinate
\(\Lambda_h\), and no further boundary contribution with the same quantum
numbers is present.  Under these assumptions the one-instanton sector defines
a holomorphic chiral \(F\)-term up to an overall nonzero constant depending on
the normalization of \(\Lambda_h\).
\end{hypothesis}

\paragraph{Higgs-patch collective-coordinate calculation.}
\label{par:ads-higgs-patch-collective-coordinate-calculation}
Let \(N_f=N_c-1\).  On the dense maximal-rank meson patch one may choose
local complex coordinates
\begin{equation}
  Q^{ai}=v_i\delta^{ai},\qquad
  \widetilde Q_{ia}=\widetilde v_i\delta_{ia},
  \qquad i=1,\ldots,N_c-1,
  \label{eq:ads-diagonal-higgs-patch}
\end{equation}
with all \(v_i\widetilde v_i\neq0\), so that
\[
  M^i{}_j=\widetilde v_i v_i\,\delta^i{}_j,\qquad
  \det M=\prod_{i=1}^{N_c-1}\widetilde v_i v_i .
\]
At such a point the gauge group \(SU(N_c)\) is completely Higgsed.  A
charge-one instanton has \(4N_c\) bosonic collective coordinates, \(2N_c\)
adjoint gaugino zero modes, and \(2N_c-2\) matter-fermion zero modes.  The
Yukawa coupling to the Higgs expectation values lifts \(2N_c-2\) adjoint
zero modes together with all \(2N_c-2\) matter zero modes, leaving exactly
two gaugino zero modes for the \(d^2\theta\) superpotential measure.

The diagonal form \eqref{eq:ads-diagonal-higgs-patch} is only a local
coordinate choice on the open set \(\det M\neq0\); the final answer must be
rewritten in flavor-covariant variables.  If \(g\in SU(N_c)\) preserves all
nonzero columns \(v_ie_i\), \(i=1,\ldots,N_c-1\), then \(g e_i=e_i\) for
those \(i\).  The determinant-one condition then fixes the last basis vector
as well, so the continuous stabilizer is trivial.

For \(N_c\geq3\), the charge-one \(SU(N_c)\) instanton bosonic moduli may be
counted as four translations, one size coordinate, and the gauge-orientation
quotient
\[
  \frac{SU(N_c)}{SU(N_c-2)\times U(1)} .
\]
The orientation quotient has real dimension
\[
  (N_c^2-1)-\bigl((N_c-2)^2-1\bigr)-1
  =
  4N_c-5 .
\]
For \(N_c=2\) the orientation space is the full \(SU(2)\), which has the
same dimension \(3=4N_c-5\).  Thus the formula below holds uniformly.
The total bosonic dimension is therefore
\[
  4+1+(4N_c-5)=4N_c .
\]

The fermion zero-mode counts are the index-theorem counts in the BPST
normalization used earlier.  For a charge-one instanton,
\[
  n_{\lambda}=2T(\operatorname{adj})=2N_c,
  \qquad
  n_{\psi_Q}=N_f,\qquad
  n_{\psi_{\widetilde Q}}=N_f,
\]
because \(2T(\mathbf N_c)=2T(\overline{\mathbf N}_c)=1\).  Hence the matter
zero-mode count is \(2N_f=2N_c-2\).

The Euclidean Yukawa terms contain the finite-dimensional zero-mode
couplings obtained by inserting the scalar expectation values into the
gauge-matter interaction
\begin{equation}
  \sqrt2\,\widetilde Q^\dagger\lambda\psi_{\widetilde Q}
  -
  \sqrt2\,\psi_Q\lambda Q^\dagger
  +\text{conjugate terms},
  \label{eq:ads-yukawa-lifting-zero-mode-coupling}
\end{equation}
with spinor and color contractions inherited from
Chapter~\ref{ch:susy-gauge-theory}.  In the diagonal patch each nonzero
\(v_i\) and \(\widetilde v_i\) supplies a nondegenerate pairing between a
matter zero mode and one non-Goldstino gaugino mode.  Thus the lifting matrix
has rank \(2N_f=2N_c-2\).  The two gaugino modes generated by the broken
supersymmetries are not lifted; they are the exact Grassmann coordinates
whose integration is the \(d^2\theta\) measure.

\paragraph{Yukawa-lifting Berezin determinant on the Higgs patch.}
\label{par:ads-yukawa-lifting-berezin-determinant}
Assume the setup of the Higgs-patch collective-coordinate calculation above.
After choosing bases for the \(2N_f\) matter zero modes and for the \(2N_f\)
non-Goldstino adjoint zero modes, the zero-mode part of
\eqref{eq:ads-yukawa-lifting-zero-mode-coupling} has the finite-dimensional
form
\begin{equation}
  S_{\mathrm{lift}}
  =
  \sum_{i=1}^{N_f}
  \left(
    a_i\,\chi_i\zeta_i
    +
    b_i\,\widetilde\chi_i\widetilde\zeta_i
  \right),
  \qquad
  a_i=c_i\,\overline{v_i},\quad
  b_i=\widetilde c_i\,\overline{\widetilde v_i},
  \label{eq:ads-yukawa-lifting-diagonal-matrix}
\end{equation}
where \(c_i,\widetilde c_i\) are nonzero constants depending only on
normalizations of zero-mode wavefunctions.  With the orientation convention
and these nonzero normalization constants absorbed into a nonzero constant
\(C_{\mathrm{or}}\), the lifted zero-mode integral is
\begin{equation}
  \int
  \prod_{i=1}^{N_f}
  d\chi_i\,d\zeta_i\,d\widetilde\chi_i\,d\widetilde\zeta_i\,
  \exp(-S_{\mathrm{lift}})
  =
  C_{\mathrm{or}}
  \prod_{i=1}^{N_f}
  \overline{v_i\,\widetilde v_i}.
  \label{eq:ads-yukawa-lifting-berezin-determinant}
\end{equation}
Thus the Yukawa matrix has maximal rank precisely on the patch
\(\det M\neq0\), saturates all lifted Grassmann coordinates, and leaves only
the two exact gaugino modes.  The anti-holomorphic product in
\eqref{eq:ads-yukawa-lifting-berezin-determinant} is a statement about this
finite-dimensional Euclidean lifting determinant.  The holomorphic ADS
denominator also uses the full regulated instanton measure, the bosonic Higgs
cutoff, and the hypothesis that the sector defines a holomorphic \(F\)-term;
these data determine the reduced factor
\(\mathcal F(M)\) below.
The diagonal Higgs background separates the flavor-color directions.  The
terms in \eqref{eq:ads-yukawa-lifting-zero-mode-coupling} pair the \(Q\)
matter zero mode in the \(i\)-th direction with one lifted adjoint mode, with
coefficient proportional to \(\overline{v_i}\), and pair the
\(\widetilde Q\) matter zero mode with another lifted adjoint mode, with
coefficient proportional to \(\overline{\widetilde v_i}\).  Since
\(v_i\widetilde v_i\neq0\) on the patch, each \(2\times2\) Grassmann block
has nonzero coefficient.

For a single block,
\[
  \exp(-a\chi\zeta)=1-a\chi\zeta .
\]
The Berezin integral over \(d\chi\,d\zeta\) extracts the coefficient of the
top monomial, up to the chosen orientation sign.  Hence one block contributes
a nonzero multiple of \(a\), and the tensor product of all \(Q\) and
\(\widetilde Q\) blocks contributes
\[
  C_{\mathrm{or}}\prod_i a_i b_i
  =
  C'_{\mathrm{or}}
  \prod_i \overline{v_i\,\widetilde v_i}.
\]
Renaming this nonzero constant as \(C_{\mathrm{or}}\) gives
\eqref{eq:ads-yukawa-lifting-berezin-determinant}.
If any lifted variable is omitted, or if one of the coefficients is zero,
the top monomial in the corresponding Grassmann variables is absent and the
integral vanishes.  This proves both the saturation statement and the
maximal-rank criterion.  The two Goldstino coordinates do not appear in
\(S_{\mathrm{lift}}\), so they remain as the \(d^2\theta\) measure.

\paragraph{Radial Higgs cutoff in the ADS instanton patch.}
Keep the \(N_f=N_c-1\) Higgs-patch setup above, and write
\[
  H(v,\widetilde v)
  =
  \sum_{i=1}^{N_f}
  \bigl(|v_i|^2+|\widetilde v_i|^2\bigr).
\]
On the maximal-rank patch \(\det M\neq0\), one has
\(H(v,\widetilde v)>0\).  Suppose that a fixed angular and zero-mode term in
the regulated charge-one collective-coordinate integral has radial factor
\begin{equation}
  \rho^{2p-1}\exp(-c\,\rho^2H(v,\widetilde v))\,d\rho,
  \qquad p\in\mathbb Z_{>0},\quad c>0,
  \label{eq:ads-radial-higgs-model-integral}
\end{equation}
where \(\rho>0\) is the instanton size.  Then the radial integral is
\begin{equation}
  I_p(H)
  =
  \int_0^\infty
  \rho^{2p-1}\exp(-cH\rho^2)\,d\rho
  =
  \frac{(p-1)!}{2(cH)^p}.
  \label{eq:ads-radial-higgs-integral-value}
\end{equation}
In particular, the Higgs expectation values give exponential convergence at
large \(\rho\) and a real homogeneous factor of degree \(-2p\) under
\((v_i,\widetilde v_i)\mapsto t(v_i,\widetilde v_i)\).  This real cutoff
estimate supplies convergence and real homogeneity.  The holomorphic ADS
denominator additionally requires the Yukawa lifting rank on the boundary
\(\det M=0\) and the holomorphic invariant \(1/\det M\), supplied by the
Yukawa-lifting calculation above and
Lemma~\ref{lem:ads-maximal-rank-meson-invariant}.

If \(\det M\neq0\) in the diagonal patch, then
\(\widetilde v_i v_i\neq0\) for every \(i\), hence each summand in \(H\) is
strictly positive and \(H>0\).  For \(\rho\ge1\), the radial integrand is
bounded by \(\rho^{2p-1}\ee^{-(cH)\rho^2}\), and this function is integrable
on \([1,\infty)\) because the exponential decay dominates every polynomial.
Near \(\rho=0\), the integrand is \(\rho^{2p-1}(1+O(\rho^2))\,d\rho\), which
is integrable because \(p\ge1\).  Thus the model radial integral is finite.

Set \(u=cH\rho^2\).  Then
\[
  \rho^{2p-1}\,d\rho
  =
  \frac{u^{p-1}}{2(cH)^p}\,du .
\]
Therefore
\[
  I_p(H)
  =
  \frac{1}{2(cH)^p}
  \int_0^\infty u^{p-1}e^{-u}\,du
  =
  \frac{\Gamma(p)}{2(cH)^p}
  =
  \frac{(p-1)!}{2(cH)^p},
\]
because \(p\) is a positive integer.

If all Higgs coordinates are multiplied by \(t\), then \(H\) is multiplied by
\(|t|^2\), so \(I_p\) is multiplied by \(|t|^{-2p}\).  The absolute values
show why this radial estimate is a real analytic convergence statement, not
a holomorphic transformation law.  Approaching \(\det M=0\) also changes the
rank of the finite-dimensional Yukawa lifting matrix, so the maximal-rank
instanton calculation cannot be extended across that boundary by the radial
estimate alone.

\begin{lemma}[Maximal-rank meson invariant for the ADS patch]
\label{lem:ads-maximal-rank-meson-invariant}
Let \(n=N_c-1\), and let
\[
  U_n=\{M\in \operatorname{Mat}_{n\times n}(\mathbb C)\mid \det M\neq0\}.
\]
Suppose \(\mathcal F:U_n\to\mathbb C\) is a single-valued holomorphic
function satisfying
\begin{equation}
  \mathcal F(LMR^{-1})=\mathcal F(M),
  \qquad L,R\in \operatorname{SL}_n(\mathbb C),
  \label{eq:ads-special-flavor-invariance}
\end{equation}
and the homogeneous scaling law
\begin{equation}
  \mathcal F(sM)=s^{-n}\mathcal F(M),
  \qquad s\in\mathbb C^\times .
  \label{eq:ads-reduced-factor-homogeneity}
\end{equation}
Then
\[
  \mathcal F(M)=\frac{C}{\det M}
\]
for a constant \(C\).  For the \(N_f=N_c-1\) SQCD \(R\)-charge assignment,
\((\det M)^{-1}\) has \(R\)-charge \(2\).
\end{lemma}

\begin{proof}
The action in \eqref{eq:ads-special-flavor-invariance} is the complexified
special-flavor action on the meson matrix.  For
\(\delta\in\mathbb C^\times\), set
\[
  D_\delta=\operatorname{diag}(\delta,1,\ldots,1).
\]
Every \(M\in U_n\) with \(\det M=\delta\) lies in the same
\(\operatorname{SL}_n(\mathbb C)\) left orbit as \(D_\delta\), because
\[
  A=M D_\delta^{-1}
  \qquad\text{satisfies}\qquad
  \det A=(\det M)(\det D_\delta)^{-1}=1
\]
and hence \(A\in \operatorname{SL}_n(\mathbb C)\) with \(M=AD_\delta\).
Therefore
\(\mathcal F\) is constant on each determinant fiber.  Define
\[
  f(\delta)=\mathcal F(D_\delta),\qquad \delta\in\mathbb C^\times .
\]
Then
\[
  \mathcal F(M)=f(\det M).
\]

Now scale \(M\) by \(s\in\mathbb C^\times\).  Since
\(\det(sM)=s^n\det M\), the homogeneity condition gives
\[
  f(s^n\delta)=s^{-n}f(\delta).
\]
For arbitrary \(u\in\mathbb C^\times\), choose \(s\) with \(s^n=u\).  The
displayed identity becomes
\[
  f(u\delta)=u^{-1}f(\delta).
\]
Taking \(\delta=1\) gives \(f(u)=C/u\).  Thus
\(\mathcal F(M)=C/(\det M)\).

Finally, when \(N_f=N_c-1=n\),
\[
  R(M)=2\left(1-\frac{N_c}{N_c-1}\right)=-\frac{2}{n},
  \qquad
  R(\det M)=-2,
\]
so \(R((\det M)^{-1})=2\).
\end{proof}

\paragraph{One-instanton test of the ADS numerator and denominator.}
\label{par:ads-one-instanton-test}
Assume Hypothesis~\ref{hyp:sqcd-one-instanton-calculus}.  For
\(N_f=N_c-1\), the one-instanton contribution to the Wilsonian chiral
superpotential has the form
\[
  W_{k=1}
  =
  \kappa_{N_c}
  \frac{\Lambda_h^{2N_c+1}}{\det M},
  \qquad
  \kappa_{N_c}\neq0
\]
in the holomorphic scale convention used in the instanton measure.  This is
the \(N_f=N_c-1\) specialization of \(W_{\mathrm{ADS}}\).

For \(N_f=N_c-1\), the Higgs-patch collective-coordinate calculation above and
Hypothesis~\ref{hyp:sqcd-one-instanton-calculus} reduce the charge-one sector
to a finite-dimensional zero-mode integral of the form
\begin{equation}
  \int d^4x_0\,d^2\eta\,
  \Lambda_h^{2N_c+1}\,\mathcal F(M),
  \label{eq:ads-reduced-instanton-integral}
\end{equation}
where \(x_0\) is the instanton position, \(\eta_\alpha\) are the two exact
gaugino modes, and \(\mathcal F(M)\) is holomorphic on the patch
\(\det M\neq0\).  The one-loop holomorphic exponent is
\[
  b_0=3N_c-N_f=2N_c+1,
\]
so a charge-one instanton carries the chiral scale factor
\(\Lambda_h^{2N_c+1}\).  The Higgs expectation values cut off the large-size
region of the instanton integral through a positive term of the form
\[
  \Delta S_{\mathrm{cl}}
  =
  c\,\rho^2
  \sum_{i=1}^{N_c-1}\bigl(|v_i|^2+|\widetilde v_i|^2\bigr),
  \qquad c>0,
\]
in the bosonic collective-coordinate action, while the small-size endpoint is
absorbed into the holomorphic scale by the hypothesis.

The remaining dependence must be a holomorphic function of the meson matrix
\(M^i{}_j=\widetilde Q_jQ^i\), because the baryonic coordinates are absent
when \(N_f=N_c-1\).  The exact nonzero modes and lifted zero modes are
integrated out in a regulator preserving the special flavor group, so
\(\mathcal F\) obeys the complexified
\(\operatorname{SL}(N_f,\mathbb C)_L\times
\operatorname{SL}(N_f,\mathbb C)_R\) invariance used in
Lemma~\ref{lem:ads-maximal-rank-meson-invariant}.  The reduced factor
\(\mathcal F(M)\) in \eqref{eq:ads-reduced-instanton-integral} must carry
engineering dimension
\[
  3-(2N_c+1)=-(2N_c-2),
\]
so under the coordinate scaling \(M\mapsto sM\), where each meson entry has
dimension \(2\), it is homogeneous of degree \(-N_f\).  With \(N_f=N_c-1\),
Lemma~\ref{lem:ads-maximal-rank-meson-invariant} gives
\[
  \mathcal F(M)=\frac{C}{\det M}.
\]
The anomaly-free \(R\)-charge gives the independent convention check
\[
  R(M)=2\left(1-\frac{N_c}{N_c-1}\right)=-\frac{2}{N_c-1},
  \qquad
  R(\det M)=-2,
\]
so \((\det M)^{-1}\) has \(R\)-charge \(2\), matching the charge of a
superpotential.  Thus the one-instanton sector has the displayed form, with
the nonzero coefficient fixed by the chosen normalization of the regulated
collective-coordinate measure.

For the smallest example, \(SU(2)\) with one flavor, the calculation gives
eight bosonic collective coordinates, four gaugino zero modes, and two matter
zero modes.  The Higgs expectation value lifts two gaugino modes together
with the two matter modes, leaving \(d^2\theta\).  The formula becomes
\[
  W_{SU(2),\,N_f=1}
  =
  \kappa_2\frac{\Lambda_h^5}{M},
  \qquad M=\widetilde Q_aQ^a .
\]
This low-rank case is often the cleanest way to audit the convention: the
power \(5\) is \(3N_c-N_f\), the denominator has dimension \(2\), and
\(R(M)=-2\), so \(M^{-1}\) supplies the required superpotential
\(R\)-charge \(2\).

\paragraph{Holomorphic scale matching across a massive threshold.}
Give the \(N_f\)-flavor theory a mass \(m\) for the last flavor.  If the
low-energy \((N_f-1)\)-flavor Wilsonian theory is reached by integrating out
that massive chiral pair, then the holomorphic scales obey
\[
  \Lambda_{N_f-1}^{3N_c-(N_f-1)}
  =
  m\,\Lambda_{N_f}^{3N_c-N_f}.
\]

Match the holomorphic gauge coordinate at the threshold \(\mu=m\).  Above the
threshold the one-loop coefficient is \(3N_c-N_f\); below it is
\(3N_c-(N_f-1)\).  The invariant combinations
\[
  \mu^{3N_c-N_f}e^{2\pi\ii\tau_{\mathrm{above}}(\mu)}
  \quad\text{and}\quad
  \mu^{3N_c-(N_f-1)}e^{2\pi\ii\tau_{\mathrm{below}}(\mu)}
  \]
are equal at \(\mu=m\) after the massive determinant is included in the
Wilsonian chiral coordinate.  Eliminating the common value of \(\tau(m)\)
gives the displayed equation.

This is a coordinate-matching calculation inside the holomorphic Wilsonian
scheme.  The dynamical assumption is the preceding sentence: the massive
pair must be integrated out in a scheme whose chiral gauge coordinate and
threshold determinant have been specified.  The displayed equation is not by
itself a proof that the infrared description exists.

\begin{proposition}[ADS decoupling recursion]
\label{prop:ads-holomorphic-decoupling-recursion}
Fix \(N_c\) and let \(1\leq n<N_c-1\).  Suppose the \((n+1)\)-flavor
Wilsonian chiral theory has, on a chosen holomorphic branch of the
maximal-rank meson patch, the ADS superpotential
\[
  W_{n+1}
  =
  C_{n+1}
  \left(
    \frac{\Lambda_{+}^{3N_c-(n+1)}}{\det \mathcal M}
  \right)^{\frac{1}{N_c-n-1}},
\]
with \(C_{n+1}\neq0\).  Add a holomorphic mass term
\[
  W_{\mathrm{mass}}=m\,\mathcal M^L{}_L,\qquad L=n+1,
\]
and assume the Wilsonian threshold is described by
the holomorphic scale matching above.  On the light branch
\(\mathcal M^a{}_L=\mathcal M^L{}_a=0\), \(a=1,\ldots,n\), the effective
\(n\)-flavor superpotential is
\begin{equation}
  W_n
  =
  C_n
  \left(
    \frac{\Lambda_-^{3N_c-n}}{\det M}
  \right)^{\frac{1}{N_c-n}},
  \qquad
  \Lambda_-^{3N_c-n}=m\,\Lambda_+^{3N_c-(n+1)},
  \label{eq:ads-decoupling-recursion-result}
\end{equation}
where \(M^a{}_b\) is the light meson block and, writing \(d=N_c-n\),
\begin{equation}
  C_n
  =
  d\left(\frac{C_{n+1}}{d-1}\right)^{\frac{d-1}{d}} .
  \label{eq:ads-coefficient-recursion}
\end{equation}
In the holomorphic scale normalization for which the one-instanton seed has
\(C_{N_c-1}=1\), the recursion gives the standard representative
\[
  C_n=N_c-n .
\]
\end{proposition}

\begin{proof}
The statement is local on a holomorphic branch because fractional powers are
involved.  The branch data are the usual discrete chiral choices of the
Wilsonian description and do not affect the algebraic recursion.

Set \(d=N_c-n\geq2\), split \(\mathcal M\) into the light block \(M\), the
mixed components, and
\[
  X=\mathcal M^L{}_L .
\]
On the light branch the determinant factorizes:
\[
  \det\mathcal M=X\det M .
\]
Define the holomorphic quantity
\[
  A=\frac{\Lambda_{+}^{3N_c-(n+1)}}{\det M}.
\]
The superpotential to be extremized with respect to the massive chiral
coordinate \(X\) is therefore
\begin{equation}
  W(X)
  =
  C_{n+1}\left(\frac{A}{X}\right)^{\frac1{d-1}}
  +mX .
  \label{eq:ads-recursion-one-variable-w}
\end{equation}
Introduce
\[
  Y=\left(\frac{A}{X}\right)^{\frac1{d-1}} .
\]
The \(F_X=0\) equation is
\[
  0
  =
  -\frac{C_{n+1}}{d-1}
  A^{\frac1{d-1}}X^{-\frac{d}{d-1}}
  +m .
\]
Multiplying by \(X\) gives the useful identity
\begin{equation}
  mX=\frac{C_{n+1}}{d-1}Y .
  \label{eq:ads-recursion-mx-y}
\end{equation}
The same equation also gives
\begin{equation}
  Y^d
  =
  \frac{(d-1)m}{C_{n+1}}\,A .
  \label{eq:ads-recursion-y-power}
\end{equation}
Substituting \eqref{eq:ads-recursion-mx-y} into
\eqref{eq:ads-recursion-one-variable-w} gives
\[
  W_{\mathrm{eff}}
  =
  C_{n+1}Y+mX
  =
  \frac{d}{d-1}C_{n+1}Y .
\]
Using \eqref{eq:ads-recursion-y-power}, this becomes
\[
  W_{\mathrm{eff}}
  =
  d\left(\frac{C_{n+1}}{d-1}\right)^{\frac{d-1}{d}}
  \left(
    \frac{m\Lambda_{+}^{3N_c-(n+1)}}{\det M}
  \right)^{\frac1d}.
\]
The holomorphic scale matching above turns the numerator into
\(\Lambda_-^{3N_c-n}\), proving
\eqref{eq:ads-decoupling-recursion-result} and
\eqref{eq:ads-coefficient-recursion}.

If \(C_{N_c-1}=1\), induction on \(d=N_c-n\) gives \(C_n=d\).  Indeed, if
\(C_{n+1}=d-1\), then \eqref{eq:ads-coefficient-recursion} gives
\[
  C_n
  =
  d\left(\frac{d-1}{d-1}\right)^{\frac{d-1}{d}}
  =
  d .
\]
\end{proof}

\begin{proposition}[Massive SQCD decoupling to pure SYM branches]
\label{prop:massive-sqcd-to-pure-sym-condensate-branches}
Fix \(1\leq n<N_c\), and assume the \(n\)-flavor Wilsonian chiral sector on
the maximal-rank meson patch is described by the standard ADS representative
\begin{equation}
  W_{\mathrm{ADS},n}
  =
  d\left(
    \frac{\Lambda_n^{3N_c-n}}{\det M}
  \right)^{1/d},
  \qquad d=N_c-n,
  \label{eq:massive-sqcd-ads-standard-representative}
\end{equation}
on a chosen holomorphic branch.  Add an invertible holomorphic mass matrix
\begin{equation}
  W_{\mathrm{mass}}=m_i{}^j M^i{}_j,
  \qquad \det m\neq0 .
  \label{eq:massive-sqcd-mass-matrix}
\end{equation}
Assume that integrating out the massive meson coordinates is legitimate in
the Wilsonian chiral theory and that holomorphic threshold matching to pure
SYM holds:
\begin{equation}
  \Lambda_0^{3N_c}
  =
  (\det m)\,\Lambda_n^{3N_c-n}.
  \label{eq:massive-sqcd-pure-scale-matching}
\end{equation}
Then the massive SQCD branch superpotential is
\begin{equation}
  W_{\mathrm{eff},k}
  =
  N_c\,\Lambda_0^3
  \exp\!\left(\frac{2\pi\ii k}{N_c}\right),
  \qquad k\in\mathbb Z/N_c\mathbb Z,
  \label{eq:massive-sqcd-pure-branch-superpotential}
\end{equation}
up to the same finite normalization of \(\Lambda_0\) and \(S\) used in the
pure-SYM glueball description.  The source identity gives
\begin{equation}
  \Lambda_0^{3N_c}
  \frac{\partial W_{\mathrm{eff},k}}
       {\partial \Lambda_0^{3N_c}}
  =
  \Lambda_0^3
  \exp\!\left(\frac{2\pi\ii k}{N_c}\right).
  \label{eq:massive-sqcd-pure-source-identity}
\end{equation}
Thus massive SQCD decoupling reproduces the pure-SYM condensate root
equation and its \(N_c\) branches, conditional on the ADS input and the
massive threshold assumptions.
\end{proposition}

\begin{proof}
Write
\begin{equation}
  A=\Lambda_n^{3N_c-n},
  \qquad
  Y=\left(\frac{A}{\det M}\right)^{1/d}.
  \label{eq:massive-sqcd-y-coordinate}
\end{equation}
Then \(W_{\mathrm{ADS},n}=dY\).  For a variation of an invertible meson
matrix,
\[
  \delta\log\det M=\operatorname{tr}(M^{-1}\delta M),
  \qquad
  \delta Y
  =
  -\frac{Y}{d}\operatorname{tr}(M^{-1}\delta M).
\]
Hence
\[
  \delta W_{\mathrm{ADS},n}
  =
  -Y\,\operatorname{tr}(M^{-1}\delta M).
\]
In indices, the \(F_M=0\) equation is therefore
\begin{equation}
  m_i{}^j=Y\,(M^{-1})^j{}_i .
  \label{eq:massive-sqcd-f-term-matrix}
\end{equation}
Multiplying by \(M^i{}_j\) and summing gives
\begin{equation}
  m_i{}^jM^i{}_j=nY,
  \label{eq:massive-sqcd-trace-mass-meson}
\end{equation}
because \(M^i{}_j(M^{-1})^j{}_i=n\).  Taking determinants in
\eqref{eq:massive-sqcd-f-term-matrix} gives
\begin{equation}
  \det m=\frac{Y^n}{\det M}.
  \label{eq:massive-sqcd-determinant-f-term}
\end{equation}
Since \(Y^d=A/\det M\), equations
\eqref{eq:massive-sqcd-determinant-f-term} and
\eqref{eq:massive-sqcd-y-coordinate} imply
\[
  Y^{d+n}=A\,\det m.
\]
As \(d+n=N_c\), this is
\[
  Y^{N_c}
  =
  (\det m)\Lambda_n^{3N_c-n}
  =
  \Lambda_0^{3N_c}
\]
by \eqref{eq:massive-sqcd-pure-scale-matching}.  Thus
\[
  Y_k=\Lambda_0^3\exp\!\left(\frac{2\pi\ii k}{N_c}\right),
  \qquad k\in\mathbb Z/N_c\mathbb Z,
\]
after choosing a branch of \(\Lambda_0^3\).

At a critical point,
\[
  W_{\mathrm{eff}}
  =
  dY+m_i{}^jM^i{}_j
  =
  dY+nY
  =
  N_cY,
\]
which proves \eqref{eq:massive-sqcd-pure-branch-superpotential}.  Finally,
with \(L_0=\Lambda_0^{3N_c}\), the branch value is
\[
  W_{\mathrm{eff},k}=N_c\,L_0^{1/N_c}\ee^{2\pi\ii k/N_c}.
\]
Therefore
\[
  L_0\frac{\partial W_{\mathrm{eff},k}}{\partial L_0}
  =
  L_0^{1/N_c}\ee^{2\pi\ii k/N_c},
\]
which is \eqref{eq:massive-sqcd-pure-source-identity}.  The proof is a
finite-dimensional chiral-coordinate elimination.  It does not by itself
construct the pure-SYM Hilbert space, prove the mass gap, or justify the ADS
and threshold hypotheses from a microscopic regulator.
\end{proof}

\paragraph{Mass-source and Konishi identities in massive SQCD.}
\label{par:massive-sqcd-konishi-source-identities}
Assume the hypotheses and notation of
Proposition~\ref{prop:massive-sqcd-to-pure-sym-condensate-branches}.  On a
chosen massive branch, the holomorphic source derivative with respect to the
mass matrix gives the meson expectation
\begin{equation}
  \frac{\partial W_{\mathrm{eff},k}}{\partial m_i{}^j}
  =
  \langle M^i{}_j\rangle_k .
  \label{eq:massive-sqcd-mass-source-meson}
\end{equation}
The finite-dimensional \(F_M=0\) equation then implies the matrix identity
\begin{equation}
  \langle M^a{}_j\rangle_k\,m_i{}^j
  =
  \langle S\rangle_k\,\delta_i{}^a,
  \qquad
  \langle S\rangle_k
  =
  \Lambda_0^3\exp\!\left(\frac{2\pi\ii k}{N_c}\right).
  \label{eq:massive-sqcd-konishi-matrix-identity}
\end{equation}
Taking the trace gives
\begin{equation}
  m_i{}^j\langle M^i{}_j\rangle_k
  =
  n\,\langle S\rangle_k .
  \label{eq:massive-sqcd-konishi-trace-identity}
\end{equation}
This is the massive-SQCD chiral-coordinate realization of the Konishi
relation.  The local operator identity with the Konishi current and
\(\bar D^2\)-exact terms requires the regulator-level current statement of
Chapter~\ref{ch:susy-nonrenormalization-holomorphy}; the calculation here
uses only the already stated Wilsonian source calculus.

The full chiral function before eliminating \(M\) is
\[
  W(M;m,L_n)
  =
  d\left(\frac{L_n}{\det M}\right)^{1/d}
  +m_i{}^jM^i{}_j,
  \qquad L_n=\Lambda_n^{3N_c-n}.
\]
Let \(M_k(m,L_n)\) be the critical point on the chosen branch.  The
envelope theorem for a holomorphic finite-dimensional critical point gives
\[
  \frac{\partial}{\partial m_i{}^j}
  W(M_k(m,L_n);m,L_n)
  =
  M_k{}^i{}_j ,
\]
because the terms involving
\(\partial M_k/\partial m_i{}^j\) are multiplied by \(F_M\) and vanish at
the critical point.  This proves
\eqref{eq:massive-sqcd-mass-source-meson}.

Equation~\eqref{eq:massive-sqcd-f-term-matrix} says
\[
  m_i{}^j=Y_k(M_k^{-1})^j{}_i .
\]
Multiplying by \(M_k{}^a{}_j\) gives
\[
  M_k{}^a{}_j\,m_i{}^j
  =
  Y_k\,M_k{}^a{}_j(M_k^{-1})^j{}_i
  =
  Y_k\,\delta_i{}^a .
\]
By
\eqref{eq:massive-sqcd-pure-source-identity}, the same branch coordinate is
the glueball source derivative:
\[
  \langle S\rangle_k=Y_k
  =
  \Lambda_0^3\exp\!\left(\frac{2\pi\ii k}{N_c}\right).
\]
This gives \eqref{eq:massive-sqcd-konishi-matrix-identity}.  Setting
\(a=i\) and summing over \(i=1,\ldots,n\) gives
\eqref{eq:massive-sqcd-konishi-trace-identity}.

\section{Quantum Modified Constraint and Confining Decoupling}

For \(N_f=N_c\), the classical relation between mesons and baryons is
\[
  \det M-B\widetilde B=0 .
\]
The holomorphic scale has dimension \(2N_c\) and the same spurionic quantum
numbers as the left side.  Under the Wilsonian and infrared hypotheses made
explicit below, the protected quantum chiral-coordinate ring is represented by
\[
  \det M-B\widetilde B=\Lambda_h^{2N_c}.
\]
This equation is a statement about the chiral ring and its vacuum space.  It
does not by itself determine the full Hilbert space, massive spectrum, or
metric on the moduli space.

\begin{hypothesis}[\(N_f=N_c\) and \(N_f=N_c+1\) chiral-sector descriptions]
\label{hyp:sqcd-quantum-modified-confining-input}
The following derivations compare protected chiral-coordinate rings in a
supersymmetric Wilsonian scheme.  For \(N_f=N_c\), the infrared chiral
sector is assumed to be generated by \(M,B,\widetilde B\), with no generated
superpotential and with a possible holomorphic deformation of the classical
relation.  For \(N_f=N_c+1\), the infrared chiral sector is assumed to be
generated by mesons \(M^I{}_J\), baryons \(B_I\), and antibaryons
\(\widetilde B^J\), \(I,J=1,\ldots,N_c+1\), with superpotential
\begin{equation}
  W_{N_c+1}
  =
  \frac{1}{\Lambda_{+}^{2N_c-1}}
  \left(B_I M^I{}_J\widetilde B^J-\det M\right).
  \label{eq:sqcd-nc-plus-one-confining-superpotential}
\end{equation}
Here \(\Lambda_{+}\) is the holomorphic scale of the \(N_f=N_c+1\) theory.
The statements below derive consequences of these chiral descriptions and of
holomorphic decoupling; they do not construct the full Hilbert space metric
or prove confinement from a microscopic regulator.
\end{hypothesis}

\paragraph{Uniqueness of the field-independent quantum modification.}
Assume the \(N_f=N_c\) chiral-sector part of
Hypothesis~\ref{hyp:sqcd-quantum-modified-confining-input}.  If the
classical constraint is deformed by a field-independent holomorphic scale
term, then up to the normalization of \(\Lambda_h\) the only allowed
deformation is
\[
  \det M-B\widetilde B=\Lambda_h^{2N_c}.
\]

For \(N_f=N_c\), the anomaly-free assignment gives
\[
  R(Q)=R(\widetilde Q)=0 .
\]
Thus \(R(M)=R(B)=R(\widetilde B)=0\).  The two classical terms have the same
engineering dimension:
\[
  [\det M]=2N_c,\qquad [B\widetilde B]=N_c+N_c=2N_c,
\]
and the same flavor and baryon quantum numbers.  A field-independent
holomorphic deformation must therefore have dimension \(2N_c\), zero
baryon charge, and the same spurionic flavor and \(R\)-charges.  The
holomorphic scale of the \(N_f=N_c\) theory appears in the invariant
combination
\[
  \Lambda_h^{b_0},\qquad b_0=3N_c-N_f=2N_c .
\]
No other field-independent monomial in the holomorphic scale has the required
dimension.  Absorbing the nonzero numerical coefficient into the definition
of \(\Lambda_h^{2N_c}\) gives the displayed constraint.

\paragraph{Mass decoupling from \(N_f=N_c+1\) to \(N_f=N_c\).}
Assume Hypothesis~\ref{hyp:sqcd-quantum-modified-confining-input}.  Add a
holomorphic mass term for the last flavor of the \(N_f=N_c+1\) theory,
\[
  W_{\mathrm{mass}}=m\,M^{L}{}_{L},
  \qquad L=N_c+1 .
\]
On the branch where the mixed heavy-light chiral coordinates vanish, the
\(F\)-term equations of
\[
  W_{N_c+1}+W_{\mathrm{mass}}
\]
imply the \(N_f=N_c\) quantum-modified constraint
\begin{equation}
  \det \widehat M-B\widetilde B=\Lambda_-^{2N_c},
  \qquad
  \Lambda_-^{2N_c}=m\,\Lambda_+^{2N_c-1},
  \label{eq:sqcd-qm-from-sconfining-decoupling}
\end{equation}
where \(\widehat M^a{}_b\), \(a,b=1,\ldots,N_c\), is the light meson block
and \(B=B_L,\widetilde B=\widetilde B^L\) are the light baryons.

Split the flavor indices as \(I=(a,L)\), with \(a=1,\ldots,N_c\).  On the
branch used for the low-energy \(N_f=N_c\) variables, set
\[
  M^a{}_L=M^L{}_a=0,\qquad B_a=\widetilde B^a=0,
\]
and write \(X=M^L{}_L\).  Then
\[
  \det M=X\det\widehat M,
  \qquad
  B_I M^I{}_J\widetilde B^J=B_LX\widetilde B^L .
\]
The superpotential restricted to these coordinates is
\[
  W
  =
  \frac{X}{\Lambda_+^{2N_c-1}}
  \left(B\widetilde B-\det\widehat M\right)
  +mX .
\]
The \(F\)-term equation for \(X\) gives
\[
  \frac{B\widetilde B-\det\widehat M}{\Lambda_+^{2N_c-1}}+m=0,
\]
or
\[
  \det\widehat M-B\widetilde B=m\,\Lambda_+^{2N_c-1}.
\]
The remaining \(F\)-term equations along the low-energy branch set
\[
  X=0
\]
unless one sits on special subloci where additional massive coordinates have
not yet been eliminated.  Finally, holomorphic decoupling at the mass
threshold gives
\[
  \Lambda_-^{3N_c-N_c}
  =
  m\,\Lambda_+^{3N_c-(N_c+1)}
  =
  m\,\Lambda_+^{2N_c-1},
\]
which is exactly \eqref{eq:sqcd-qm-from-sconfining-decoupling}.

\begin{remark}[Checks on the \(N_f=N_c+1\) confining superpotential]
\label{rem:sqcd-sconfining-superpotential-checks}
For \(N_f=N_c+1\), the two terms in
\eqref{eq:sqcd-nc-plus-one-confining-superpotential} have the same
engineering dimension and \(R\)-charge, and the prefactor makes
\(W_{N_c+1}\) a dimension-three, \(R\)-charge-two chiral superpotential.
At \(N_f=N_c+1\),
\[
  R(Q)=R(\widetilde Q)=\frac1{N_c+1}.
\]
Thus
\[
  R(M)=\frac2{N_c+1},\qquad
  R(B_I)=R(\widetilde B^J)=\frac{N_c}{N_c+1}.
\]
The product \(B_IM^I{}_J\widetilde B^J\) has \(R\)-charge
\[
  \frac{N_c}{N_c+1}+\frac2{N_c+1}
  +\frac{N_c}{N_c+1}
  =
  2 ,
\]
and \(\det M\), where the determinant is over the \(N_f=N_c+1\) flavor
indices, has \(R\)-charge
\[
  (N_c+1)\frac2{N_c+1}=2 .
\]
Classically, \(\dim M=2\) and \(\dim B_I=\dim\widetilde B^J=N_c\).  Both
\(B_IM^I{}_J\widetilde B^J\) and \(\det M\) therefore have dimension
\[
  2N_c+2=2(N_c+1).
\]
Since \(\Lambda_+^{2N_c-1}\) has dimension \(2N_c-1\) and zero
\(R\)-charge in the anomaly-free spurion convention, the quotient has
dimension \(3\) and \(R\)-charge \(2\), as required.
\end{remark}

\section{Superconformal SQCD Bookkeeping}
\label{sec:n1-superconformal-sqcd-bookkeeping}

The same anomaly calculation also fixes the candidate superconformal data of SQCD
in the range where an interacting infrared fixed point exists.  For a chiral
primary operator \(\mathcal O\) in a four-dimensional \(\mathcal N=1\)
superconformal theory,
\[
  \Delta(\mathcal O)=\frac32 R(\mathcal O).
\]
For an elementary chiral field \(\Phi\), write its scalar scaling dimension
as
\[
  \Delta(\Phi)=1+\frac12\gamma_\Phi .
\]
Compatibility of these two conventions gives
\begin{equation}
  \gamma_\Phi=3R(\Phi)-2 .
  \label{eq:n1-gamma-from-r}
\end{equation}
Here \(\gamma_\Phi\) is the chapter convention
\[
  \gamma_\Phi=\frac{d}{d\log\mu}\log Z_\Phi .
\]
Some supersymmetric-gauge-theory references instead use
\(\mathcal C_\Phi=\frac12\,d\log Z_\Phi/d\log\mu\).  The conversion is
\[
  \gamma_\Phi=2\mathcal C_\Phi,\qquad
  \Delta(\Phi)=1+\mathcal C_\Phi .
\]
For a product of two chiral fields with equal anomalous dimensions, the
product anomalous dimension in the \(\mathcal C\)-notation is numerically
equal to the single-field \(\gamma_\Phi\) used below:
\[
  \mathcal C_{\Phi\Psi}
  =
  \Delta(\Phi\Psi)-2
  =
  \mathcal C_\Phi+\mathcal C_\Psi
  =
  \gamma_\Phi
  \qquad
  (\gamma_\Phi=\gamma_\Psi).
\]
This is why the conifold formulas may be written with \(1-\gamma_A\) here
while the same fixed point is sometimes described by
\(\mathcal C_{AB}=-\frac12\) for the meson \(A_iB_j\).

For SQCD with no tree superpotential, the anomaly-free assignment of the
previous section gives
\[
  R(Q)=R(\widetilde Q)=1-\frac{N_c}{N_f},
  \qquad
  \gamma_Q=1-\frac{3N_c}{N_f}.
\]
Substitution in the NSVZ numerator gives
\[
  3N_c-N_f(1-\gamma_Q)
  =
  3N_c-N_f\left(\frac{3N_c}{N_f}\right)
  =
  0 .
\]
Thus the anomaly-free \(R\)-charge is also the unique candidate
superconformal \(R\)-charge compatible with the flavor-symmetric NSVZ
coordinate.

The meson \(M=\widetilde Q Q\) has
\[
  R(M)=2\left(1-\frac{N_c}{N_f}\right),
  \qquad
  \Delta(M)=3\left(1-\frac{N_c}{N_f}\right).
\]
The scalar unitarity bound for a gauge-invariant chiral operator requires
\(\Delta(M)\geq1\), which gives \(N_f\geq\frac32N_c\).  This is the lower
edge of the usual electric conformal-window test; below it the electric
meson cannot be an interacting chiral primary with the displayed
superconformal \(R\)-charge.  The upper edge \(N_f<3N_c\) is the condition
that the electric gauge coupling is asymptotically free in the holomorphic
one-loop coordinate.  The existence and uniqueness of the infrared fixed
point in the full interval require the nonperturbative input summarized in
the status boundary below.

\paragraph{Electric Gaussian endpoint from the first nonzero beta term.}
At \(N_f=3N_c\), the one-loop coefficient \(b_0=3N_c-N_f\) vanishes, so the
direction of the perturbative flow is not determined by \(b_0\) itself.  In
the chapter's canonical coupling coordinate the weak-coupling matter
anomalous dimension for SQCD with \(W=0\) is
\begin{equation}
  \gamma_Q(g)
  =
  -\frac{C_2(\mathbf N_c)}{8\pi^2}g^2+O(g^4),
  \qquad
  C_2(\mathbf N_c)=\frac{N_c^2-1}{2N_c}.
  \label{eq:sqcd-one-loop-quark-gamma}
\end{equation}
Substituting this into the specialized NSVZ expression
\eqref{eq:sqcd-nsvz-beta-specialized} gives, at \(N_f=3N_c\),
\begin{equation}
  \beta_g
  =
  \frac{3N_cC_2(\mathbf N_c)}{128\pi^4}g^5+O(g^7)
  =
  \frac{3(N_c^2-1)}{256\pi^4}g^5+O(g^7).
  \label{eq:sqcd-electric-gaussian-edge-beta}
\end{equation}
Thus \(g=0\) is marginally irrelevant toward the infrared: lowering the
renormalization scale drives a sufficiently small electric coupling back to
zero.  This is the dynamical perturbative reason for calling \(N_f=3N_c\)
the free electric Gaussian edge.  The equality of
\eqref{eq:sqcd-free-field-a-comparison} with the free-field value at the same
point is only kinematic anomaly arithmetic; it does not by itself determine
the RG direction.

\begin{hypothesis}[SCFT input used in bookkeeping]
\label{hyp:n1-scft-bookkeeping-input}
Whenever this chapter assigns a scaling dimension by
\(\Delta=\frac32R\), the following assumptions are in force: the infrared
limit exists as a unitary four-dimensional \(\mathcal N=1\) SCFT; the
operator under discussion is a chiral primary in that SCFT; the displayed
\(U(1)_R\) current is the superconformal \(R\)-current after all allowed
mixing with non-\(R\) flavor currents has been fixed; and accidental
symmetries do not change the \(R\)-charge of the operator in question.  The
anomaly and NSVZ calculations below test these assumptions; they do not by
themselves construct the fixed point.
\end{hypothesis}

\begin{quotedtheorem}[\(a\)-maximization status boundary]
\label{qt:n1-a-maximization-boundary}
Use the four-dimensional \(\mathcal N=1\) \(a\)-maximization theorem of
\cite{IntriligatorWecht2003}, together with the anomaly/current two-point
normalizations of \cite{AnselmiFreedmanGrisaruJohansen1998}.  Assume a
unitary four-dimensional \(\mathcal N=1\) SCFT with a stress-tensor
supermultiplet, a finite declared basis of conserved non-\(R\) flavor currents
\(F_I\), and anomaly coefficients computed by matching to any ultraviolet
description or dual description with the same global-current background
fields.  Let
\[
  R_t=R_0+\sum_I s^I F_I
\]
range over gauge-anomaly-free and superpotential-compatible trial
\(R\)-currents in this basis.  The trial anomaly functional is
\begin{equation}
  a_{\rm trial}(s)
  =
  \frac3{32}
  \left(3\operatorname{Tr}R_t^3-\operatorname{Tr}R_t\right),
  \label{eq:n1-a-trial-functional}
\end{equation}
where the traces are over Weyl fermions.  If the exact superconformal
\(R\)-current lies in the declared trial space and no accidental symmetry has
been omitted, it is a local maximum of
\eqref{eq:n1-a-trial-functional}.  Its stationarity equations are
\begin{equation}
  0=
  \frac{\partial a_{\rm trial}}{\partial s^I}
  =
  \frac3{32}
  \left(9\operatorname{Tr}R_\ast^2F_I-\operatorname{Tr}F_I\right),
  \label{eq:n1-a-max-stationarity}
\end{equation}
and the Hessian at the solution is
\begin{equation}
  \frac{\partial^2 a_{\rm trial}}{\partial s^I\partial s^J}
  =
  \frac{27}{16}\operatorname{Tr}R_\ast F_I F_J
  =
  -\frac9{16}\tau_{IJ}.
  \label{eq:n1-a-max-hessian-current-two-point}
\end{equation}
Here \(\tau_{IJ}=-3\operatorname{Tr}R_\ast F_I F_J\) is the positive
flavor-current two-point matrix in the present normalization.  Positivity of
\(\tau\) is the physical input that turns stationarity into a local maximum.
\end{quotedtheorem}

The theorem applies after the SCFT and the trial-current space have been
specified.  Before applying it one must identify all conserved Abelian flavor
currents that are allowed to mix with \(R\) in the symmetry sector being
studied.  Cartan generators of a non-Abelian flavor group may be included only
if the proposed fixed point or deformation has broken the non-Abelian symmetry
to that Abelian subgroup; otherwise mixing with such a Cartan would choose a
direction in the multiplet and is not in the preserved-symmetry trial space.
If a gauge-invariant chiral operator reaches the scalar unitarity bound
\[
  \Delta\ge1,\qquad R\ge\frac23,
\]
it becomes a free chiral multiplet and brings an accidental current.  Then
the trial space and the anomaly functional must be enlarged before
maximization.

\begin{example}[Unitarity-bound correction for an accidental chiral field]
\label{ex:n1-accidental-symmetry-a-correction}
For one chiral multiplet of trial superfield charge \(r\), the contribution to
\eqref{eq:n1-a-trial-functional} is
\begin{equation}
  a_\chi(r)
  =
  \frac3{32}\left(3(r-1)^3-(r-1)\right),
  \qquad
  a_\chi\!\left(\frac23\right)=\frac1{48}.
  \label{eq:chiral-multiplet-a-contribution}
\end{equation}
If a gauge-invariant chiral operator \(M\), with multiplicity \(d_M\), would
have naive trial charge \(R_M(s)<2/3\), the old anomaly polynomial is no
longer the object to maximize.  The minimal decoupled-field correction is
\begin{equation}
  a_{\rm corrected}(s)
  =
  a_{\rm naive}(s)
  +
  d_M\left[
    a_\chi\!\left(\frac23\right)-a_\chi(R_M(s))
  \right],
  \label{eq:a-max-accidental-correction}
\end{equation}
with an additional accidental \(U(1)_M\) acting on the free field.  In SQCD,
the naive electric meson has
\[
  R(M)=2\left(1-\frac{N_c}{N_f}\right).
\]
For \(N_f<\frac32N_c\) this is below \(2/3\).  The failure is not a small
arithmetic defect in \(a_{\rm naive}\); it is evidence that the assumed
interacting chiral-primary spectrum and current basis are incomplete.  This is
why the conformal-window bookkeeping above stops at the meson unitarity test
unless the magnetic/free-field description and accidental currents are made
part of the infrared datum.
\end{example}

\paragraph{SQCD central charges and free-field comparison.}
Assume Hypothesis~\ref{hyp:n1-scft-bookkeeping-input} and suppose
\(\frac32N_c<N_f<3N_c\).  The current-identification boundary is the same as
in Quoted Theorem~\ref{qt:n1-a-maximization-boundary}: possible mixing with
conserved flavor currents must already have been fixed in the declared
flavor-symmetric sector, and the meson unitarity-bound failure mode is absent
in this range.  For the candidate SQCD superconformal \(R\)-current displayed
above,
\begin{equation}
  \operatorname{Tr}R=-N_c^2-1,
  \qquad
  \operatorname{Tr}R^3
  =
  N_c^2-1-\frac{2N_c^4}{N_f^2}.
  \label{eq:sqcd-superconformal-r-traces}
\end{equation}
Consequently the conformal anomalies are
\begin{equation}
  a_{\mathrm{IR}}
  =
  \frac3{32}
  \left(
    4N_c^2-2-\frac{6N_c^4}{N_f^2}
  \right),
  \qquad
  c_{\mathrm{IR}}
  =
  \frac1{32}
  \left(
    14N_c^2-4-\frac{18N_c^4}{N_f^2}
  \right).
  \label{eq:sqcd-central-charges}
\end{equation}
The free ultraviolet theory has
\begin{equation}
  a_{\mathrm{UV}}
  =
  \frac3{16}(N_c^2-1)+\frac{N_cN_f}{24}.
  \label{eq:sqcd-free-uv-a}
\end{equation}
Writing \(y=N_f/N_c\), the difference is
\begin{equation}
  a_{\mathrm{UV}}-a_{\mathrm{IR}}
  =
  \frac{N_c^2}{48y^2}
  (3-y)^2(2y+3).
  \label{eq:sqcd-free-field-a-comparison}
\end{equation}
It is strictly positive in the interacting conformal window and vanishes at
the Gaussian edge \(N_f=3N_c\).  This is only an anomaly-arithmetic
comparison between a free ultraviolet fixed point and the candidate
infrared SCFT.  It is not a proof of a monotonicity theorem or of the
existence of the flow; those remain part of the independent continuum and
SCFT hypotheses.

The traces are over Weyl fermions.  The gaugino contributes
\[
  \operatorname{Tr}R_{\lambda}=N_c^2-1,
  \qquad
  \operatorname{Tr}R_{\lambda}^3=N_c^2-1 .
\]
The scalar \(R\)-charge of \(Q\) and \(\widetilde Q\) is
\[
  R(Q)=R(\widetilde Q)=1-\frac{N_c}{N_f},
\]
so their fermions have \(R\)-charge \(-N_c/N_f\).  There are
\(2N_cN_f\) such Weyl fermions.  Hence
\[
  \operatorname{Tr}R
  =
  (N_c^2-1)+2N_cN_f\left(-\frac{N_c}{N_f}\right)
  =
  -N_c^2-1,
\]
and
\[
  \operatorname{Tr}R^3
  =
  (N_c^2-1)
  +2N_cN_f\left(-\frac{N_c}{N_f}\right)^3
  =
  N_c^2-1-\frac{2N_c^4}{N_f^2}.
\]
For a four-dimensional \(\mathcal N=1\) SCFT,
\begin{equation}
  a=\frac3{32}(3\operatorname{Tr}R^3-\operatorname{Tr}R),
  \qquad
  c=\frac1{32}(9\operatorname{Tr}R^3-5\operatorname{Tr}R).
  \label{eq:n1-central-charge-from-r-traces}
\end{equation}
Substituting the two traces gives \eqref{eq:sqcd-central-charges}.

At the ultraviolet Gaussian fixed point, a vector multiplet contributes
\(\frac3{16}\dim SU(N_c)\) to \(a\), and a free chiral multiplet contributes
\(\frac1{48}\) per complex gauge-flavor component.  Since
\(\dim SU(N_c)=N_c^2-1\) and there are \(2N_cN_f\) chiral components,
\eqref{eq:sqcd-free-uv-a} follows.

Subtracting the candidate infrared value from the ultraviolet value gives
\[
  a_{\mathrm{UV}}-a_{\mathrm{IR}}
  =
  -\frac{3N_c^2}{16}
  +\frac{N_cN_f}{24}
  +\frac{9N_c^4}{16N_f^2}.
\]
Set \(y=N_f/N_c\).  Then
\[
  a_{\mathrm{UV}}-a_{\mathrm{IR}}
  =
  N_c^2
  \left(
    -\frac3{16}+\frac{y}{24}+\frac{9}{16y^2}
  \right)
  =
  \frac{N_c^2}{48y^2}\bigl(2y^3-9y^2+27\bigr).
\]
The polynomial factorizes as
\[
  2y^3-9y^2+27=(3-y)^2(2y+3).
\]
For \(y>0\) this expression is nonnegative and it is zero only when \(y=3\).
In the interacting window \(3/2<y<3\), it is strictly positive.  The proof
has established this finite anomaly comparison; it has not established the
general \(a\)-theorem.

\section{Seiberg Duality And The SQCD Phase Structure}
\label{sec:n1-seiberg-duality-sqcd}

The preceding superconformal bookkeeping used only the electric SQCD
variables.  We now isolate the nonperturbative duality input that is needed
whenever the chapter talks about the full SQCD phase diagram.  This is a
field-theoretic statement about four-dimensional \(\mathcal N=1\) gauge
dynamics; no string construction is part of the definition or the argument.

For charge tables it is useful to display the current algebra as
\[
  SU(N_f)_L\times SU(N_f)_R\times U(1)_B\times U(1)_R .
\]
That line is not yet the compact group that can be coupled to background
bundles.  With the normalization used in this chapter, \(q_B(Q)=+1\) and
\(q_B(\widetilde Q)=-1\), let
\[
  \zeta_f=e^{2\pi i/N_f},
  \qquad
  \omega_c=e^{2\pi i/N_c}.
\]
The transformations
\[
  (\zeta_f\mathbf 1_L,\zeta_f\mathbf 1_R,e^{-2\pi i/N_f})
  \quad\text{and}\quad
  (\mathbf 1_L,\mathbf 1_R,e^{2\pi i/N_c})
\]
act trivially on electric gauge-invariant local operators: the first acts
trivially already on \(Q,\widetilde Q\), while the second acts as the
\(SU(N_c)\) gauge-center element \(\omega_c\) on \(Q\) and \(\omega_c^{-1}\)
on \(\widetilde Q\).  Conversely, if
\((\zeta_f^a\mathbf 1_L,\zeta_f^b\mathbf 1_R,e^{i\beta})\) is invisible on
local gauge-invariant operators, then on the elementary quarks it can differ
only by a color-center transformation.  Thus for some \(s\in\mathbb Z/N_c\),
\begin{equation}
  \zeta_f^a e^{i\beta}=\omega_c^s,
  \qquad
  \zeta_f^{-b} e^{-i\beta}=\omega_c^{-s}.
  \label{eq:sqcd-faithful-kernel-equations}
\end{equation}
Multiplying the two equations gives \(a=b\) modulo \(N_f\), and then
\(e^{i\beta}=\omega_c^s\zeta_f^{-a}\).  Hence the faithful compact
flavor-baryon group acting on local gauge-invariant operators is
\begin{equation}
  G_{FB}^{\rm SQCD}
  =
  \frac{SU(N_f)_L\times SU(N_f)_R\times U(1)_B}
  {\mathbb Z_{N_f}^{\rm diag}\times\mathbb Z_{N_c}^{B/{\rm gauge}}}.
  \label{eq:sqcd-faithful-flavor-baryon-group}
\end{equation}
The anomaly-free \(R\)-current below is used first as a Lie-algebra current.
Since \(R(Q)=R(\widetilde Q)=1-N_c/N_f\) is fractional in the conventional
normalization, a compact \(U(1)_R\) background requires an integral
normalization, for example \(\widehat R=N_fR\), together with the appropriate
spin/\(R\)-structure data.  The local anomaly-polynomial calculation below
does not by itself compare those torsion choices.

An admissible flavor-baryon background on a spacetime \(X\) is therefore a
principal \(G_{FB}^{\rm SQCD}\)-bundle, not necessarily a product of separate
\(SU(N_f)_L\), \(SU(N_f)_R\), and \(U(1)_B\) bundles.  On a good cover it may
be represented by local lifts
\((g^L_{\alpha\beta},g^R_{\alpha\beta},e^{i\beta_{\alpha\beta}})\) whose
triple-overlap products are
\[
  g^L_{\alpha\beta}g^L_{\beta\gamma}g^L_{\gamma\alpha}
  =
  \zeta_f^{a_{\alpha\beta\gamma}}\mathbf 1,
  \qquad
  g^R_{\alpha\beta}g^R_{\beta\gamma}g^R_{\gamma\alpha}
  =
  \zeta_f^{a_{\alpha\beta\gamma}}\mathbf 1,
\]
\begin{equation}
  e^{i(\beta_{\alpha\beta}+\beta_{\beta\gamma}
  +\beta_{\gamma\alpha})}
  =
  \zeta_f^{-a_{\alpha\beta\gamma}}\,
  \omega_c^{s_{\alpha\beta\gamma}},
  \label{eq:sqcd-admissible-background-cocycles}
\end{equation}
with \(a_{\alpha\beta\gamma}\in\mathbb Z/N_f\) and
\(s_{\alpha\beta\gamma}\in\mathbb Z/N_c\).  Electric quarks are then sections
of the combined color--flavor--baryon bundle; the color center cocycle cancels
the \(s_{\alpha\beta\gamma}\) obstruction.  The ordinary direct-product
background used in the triangle calculation is the special case
\(a=s=0\).  On general \(G_{FB}^{\rm SQCD}\)-bundles there can be additional
discrete inflow classes and quotient-dependent contact terms, so equality of
the full background anomaly is a separate part of the duality conjecture.

The electric fields transform as
\[
\begin{array}{c|c|c|c|c|c}
  & SU(N_c) & SU(N_f)_L & SU(N_f)_R & U(1)_B & U(1)_R \\ \hline
  Q^i & \mathbf N_c & \mathbf N_f & \mathbf 1
      & +1 & 1-\frac{N_c}{N_f} \\
  \widetilde Q_j & \overline{\mathbf N}_c & \mathbf 1
      & \overline{\mathbf N}_f
      & -1 & 1-\frac{N_c}{N_f}
\end{array}
\]
and the meson \(M^i{}_j=\widetilde Q_jQ^i\) transforms as
\((\mathbf N_f,\overline{\mathbf N}_f)\).  The normalization of
\(U(1)_B\) is chosen so that an electric baryon \(Q^{N_c}\) has charge
\(N_c\).  This normalization is convenient for local charge arithmetic, but
\eqref{eq:sqcd-faithful-flavor-baryon-group} shows that it is not globally
harmless when compact background fluxes are present.

For \(N_f\geq N_c+2\), set
\[
  \widetilde N_c=N_f-N_c .
\]
The magnetic SQCD variables consist of an \(SU(\widetilde N_c)\) vector
multiplet, magnetic quarks \(q_i,\widetilde q^j\), and a gauge-singlet meson
coordinate \(M^i{}_j\) with the global charges
\[
\begin{array}{c|c|c|c|c|c}
  & SU(\widetilde N_c) & SU(N_f)_L & SU(N_f)_R
  & U(1)_B & U(1)_R \\ \hline
  q_i & \mathrm{fund} & \overline{\mathbf N}_f & \mathbf 1
      & \frac{N_c}{\widetilde N_c} & \frac{N_c}{N_f} \\
  \widetilde q^j & \overline{\mathrm{fund}} & \mathbf 1 & \mathbf N_f
      & -\frac{N_c}{\widetilde N_c} & \frac{N_c}{N_f} \\
  M^i{}_j & \mathbf 1 & \mathbf N_f & \overline{\mathbf N}_f
      & 0 & 2\left(1-\frac{N_c}{N_f}\right).
\end{array}
\]
The magnetic holomorphic matching coordinate has superpotential
\begin{equation}
  W_{\mathrm{mag}}
  =
  \frac{\lambda}{\mu_*}M^i{}_j q_i\widetilde q^j .
  \label{eq:sqcd-magnetic-superpotential}
\end{equation}
Here \(M^i{}_j\) is normalized as the electric composite
\(\widetilde Q_jQ^i\), so it has engineering dimension two in the
microscopic matching convention; \(\mu_*\) is a nonzero matching scale of
dimension one, and \(\lambda\) is a dimensionless holomorphic coefficient.
The often-used convention \(\lambda=1\) fixes a holomorphic coordinate; it does
not fix the nonholomorphic metric on field space.

A canonical gauge--Yukawa beta function requires a magnetic Wilsonian action,
not just the chiral superpotential.  At a declared normalization scale \(\mu_0\)
and in a declared mass-independent canonical scheme, write the quadratic
magnetic Kähler terms as
\begin{equation}
  \int d^4\theta\left[
    \frac{Z_M(\mu_0)}{|\mu_*|^2}
    M^\dagger{}^j{}_i M^i{}_j
    +
    Z_q(\mu_0)\,q^\dagger e^{V_m}q
    +
    Z_{\widetilde q}(\mu_0)\,
    \widetilde q^\dagger e^{-V_m}\widetilde q
  \right].
  \label{eq:sqcd-magnetic-kahler-normalization}
\end{equation}
In the flavor-symmetric chart \(Z_q=Z_{\widetilde q}\), but the common factor is
still part of the full-action datum.  The dimension-one holomorphic coordinate
\(M/\mu_*\) is canonical only when the meson wavefunction factor is also fixed
to one in the chosen scheme.  In general the canonical fields at \(\mu_0\) are
\begin{equation}
  M_{\mathrm{can}}=Z_M(\mu_0)^{1/2}\frac{M}{\mu_*},\qquad
  q_{\mathrm{can}}=Z_q(\mu_0)^{1/2}q,\qquad
  \widetilde q_{\mathrm{can}}
  =Z_{\widetilde q}(\mu_0)^{1/2}\widetilde q .
  \label{eq:sqcd-magnetic-canonical-fields}
\end{equation}
Thus the canonical Yukawa coordinate is
\begin{equation}
  h_{\mathrm{can}}(\mu_0)
  =
  \lambda(\mu_0)\,
  Z_M(\mu_0)^{-1/2}
  Z_q(\mu_0)^{-1/2}
  Z_{\widetilde q}(\mu_0)^{-1/2}.
  \label{eq:sqcd-magnetic-canonical-yukawa-coupling}
\end{equation}
After this full-action normalization has been declared one may write
\begin{equation}
  W_{\mathrm{mag}}
  =
  h_{\mathrm{can}}\,
  M_{\mathrm{can}}{}^i{}_j
  q_{\mathrm{can},i}\widetilde q_{\mathrm{can}}{}^j ,
  \label{eq:sqcd-magnetic-canonical-yukawa}
\end{equation}
with a dimensionless Yukawa coordinate.  The displayed holomorphic
\(M q\widetilde q/\mu_*\) form keeps the operator map visible; the canonical
form is the one used for perturbative RG statements.  Seiberg-duality and
chiral-ring data determine the holomorphic representation content and the
allowed cubic term, but they do not determine the magnetic Kähler potential or
the conversion from holomorphic matching variables to canonical running
variables.  Changing \(Z_M\) at fixed \(\lambda,Z_q,Z_{\widetilde q}\) changes
\(h_{\mathrm{can}}\); it is not a harmless rewriting of \(M/\mu_*\).

\paragraph{Magnetic gauge--Yukawa ultraviolet status.}
Let \(t=\widetilde N_c=N_f-N_c\), let \(g_m\) be the magnetic canonical gauge
coupling, set \(h=h_{\mathrm{can}}\) in the canonical chart just declared, and
let
\[
  C_F^{(m)}=C_2(\mathbf{fund}_{SU(t)})=\frac{t^2-1}{2t},
  \qquad
  b_{0,m}=3t-N_f=2N_f-3N_c .
\]
In a mass-independent canonical scheme, the one-loop anomalous dimensions of
the magnetic fields in the flavor-symmetric gauge--Yukawa chart are
\begin{align}
  \gamma_q=\gamma_{\widetilde q}
  &=
  \frac{1}{16\pi^2}
  \left(N_f|h|^2-2C_F^{(m)}g_m^2\right)
  +O(g_m^4,g_m^2|h|^2,|h|^4),
  \label{eq:sqcd-magnetic-quark-gamma-one-loop}\\
  \gamma_M
  &=
  \frac{1}{16\pi^2}
  \left(t|h|^2\right)
  +O(g_m^2|h|^2,|h|^4).
  \label{eq:sqcd-magnetic-meson-gamma-one-loop}
\end{align}
The nonrenormalization theorem then gives
\[
  \frac{\beta_h}{h}
  =
  \frac12(\gamma_M+\gamma_q+\gamma_{\widetilde q}),
\]
so, to one loop,
\begin{align}
  \beta_{g_m^2}
  &=
  -\frac{b_{0,m}}{8\pi^2}g_m^4+\cdots,
  \label{eq:sqcd-magnetic-gauge-beta-one-loop}\\
  \beta_{|h|^2}
  &=
  \frac{|h|^2}{16\pi^2}
  \left((2N_f+t)|h|^2-4C_F^{(m)}g_m^2\right)+\cdots .
  \label{eq:sqcd-magnetic-yukawa-beta-one-loop}
\end{align}
For the ratio \(r=|h|^2/g_m^2\),
\begin{equation}
  \mu\frac{dr}{d\mu}
  =
  \frac{r g_m^2}{16\pi^2}
  \left((2N_f+t)r+2b_{0,m}-4C_F^{(m)}\right)+\cdots .
  \label{eq:sqcd-magnetic-yukawa-ratio-flow}
\end{equation}
In the free electric range \(N_f\ge 3N_c\), one has \(t\ge2N_c\), and
\[
  2b_{0,m}-4C_F^{(m)}
  =
  2(t-N_c)+\frac{2}{t}
  >
  0 .
\]
Therefore no positive nonzero weak-coupling ray makes both \(g_m\) and \(h\)
asymptotically free in this range.  The line \(h=0\) is asymptotically free
for the gauge sector, but it removes the superpotential that implements the
Seiberg magnetic field identifications.  A magnetic description with nonzero
matching \(h\) must therefore be treated as a Wilsonian theory with a finite
matching cutoff and specified \(h\), or be supplied with additional
ultraviolet completion data; magnetic gauge asymptotic freedom alone is not a
UV definition of the full magnetic gauge--Yukawa Lagrangian.

The fractional magnetic baryon charges are also shorthand for a representation
of the same faithful group \eqref{eq:sqcd-faithful-flavor-baryon-group}, not
for an independent baryon circle.  Under
\((\mathbf 1_L,\mathbf 1_R,e^{2\pi i/N_c})\), \(q\) transforms by
\(e^{2\pi i/\widetilde N_c}\), the magnetic color-center phase.  Under
\((\zeta_f\mathbf 1_L,\zeta_f\mathbf 1_R,e^{-2\pi i/N_f})\), the left
antifundamental charge of \(q\) and its baryon charge combine to
\(e^{-2\pi i/\widetilde N_c}\), again a magnetic color-center phase; the
\(\widetilde q\) phase is the inverse, and \(M\) is invariant.  Thus the
magnetic fields are well-defined on the same admissible background after the
magnetic color bundle supplies the corresponding center cocycle.

\begin{hypothesis}[SQCD infrared-comparison hypotheses]
\label{hyp:sqcd-ir-comparison-hypotheses}
Whenever this section compares the electric and magnetic SQCD descriptions,
the following hypotheses are part of the statement.
\begin{enumerate}
\item The regulator and ultraviolet status of each Lagrangian description are
  part of the datum.  In the electric asymptotically free range
  \(N_c+2\le N_f<3N_c\), the electric \(SU(N_c)\) theory is defined by a
  continuum Wilsonian limit in a regulator preserving the gauge Ward
  identities, the listed flavor symmetries, and enough supersymmetry for the
  chiral-coordinate arguments of
  Chapters~\ref{ch:susy-wilsonian-schemes-exact-dynamics} and
  \ref{ch:susy-nonrenormalization-holomorphy}.  The magnetic description is
  compared in the corresponding infrared window: as an asymptotically free
  continuum theory in the interacting conformal subrange, and as an
  infrared-free Wilsonian description with a declared matching scale and
  cutoff in the free magnetic subrange.
\item For the free electric continuation \(N_f\ge 3N_c\), the electric
  variables require extra ultraviolet data.  At \(N_f=3N_c\) the electric
  one-loop coefficient vanishes, but the first nonzero term
  \eqref{eq:sqcd-electric-gaussian-edge-beta} makes the electric coupling
  marginally irrelevant toward the infrared; this is the Gaussian edge of the
  free electric family.  For \(N_f>3N_c\), the electric
  Lagrangian is used either as a finite cutoff effective theory with cutoff
  \(\Lambda_{\rm el}\), as a theory supplied with a specified ultraviolet
  completion, or as an emergent infrared coordinate system defined through a
  magnetic Wilsonian theory with matching scale, magnetic Kähler data
  \(Z_M,Z_q,Z_{\widetilde q}\), holomorphic coefficient \(\lambda\), and the
  derived canonical Yukawa datum \(h_{\mathrm{can}}(\Lambda_{\rm mag})\).
  The one-loop analysis
  \eqref{eq:sqcd-magnetic-yukawa-ratio-flow} shows that the magnetic gauge
  coupling may be asymptotically free while the full nonzero-\(h\)
  gauge--Yukawa system is not thereby UV-completed.  In the finite cutoff
  formulation one
  first takes the thermodynamic limit at fixed \(\Lambda_{\rm el}\), then the
  infrared limit of momenta, masses, and sources with all ratios
  \(p/\Lambda_{\rm el}\) tending to zero.  A subsequent
  \(\Lambda_{\rm el}\to\infty\) statement is admitted only for observables
  whose low-energy limits are explicitly declared independent of the electric
  ultraviolet completion; in this chapter those are the protected chiral
  ring, local current-anomaly coefficients in the declared background class,
  and holomorphic mass/Higgs deformation ledgers.
\item The unbroken global currents and faithful compact flavor-baryon
  background group \eqref{eq:sqcd-faithful-flavor-baryon-group} are
  represented by the same infrared operators.  On liftable product
  backgrounds the local anomaly-polynomial coefficients must match.  On
  nonliftable \(G_{FB}^{\rm SQCD}\)-bundles and on compact \(R\)-backgrounds,
  the full inflow class, torsion terms, and allowed local counterterms must be
  specified separately before one can claim compact-background anomaly
  agreement.
\item The displayed \(U(1)_R\) current is the superconformal current in the
  interacting range after all allowed flavor-current mixing has been fixed.
  At free endpoints one must include the accidental symmetries of the free
  fields before interpreting \(\Delta=\frac32R\).
\item The chiral coordinates \(M,B,\widetilde B\) and their magnetic images
  generate the protected chiral sector being compared, after imposing
  \(F\)-term equations and after excluding extra accidental chiral generators
  not present in the stated phase.
\item The observable sector compared in this chapter consists of
  gauge-invariant local operators and their correlation functions in the
  declared background class, the protected chiral ring, and holomorphic mass
  and Higgs deformations.  Equality of line/defect categories or of all
  compact-manifold partition functions is not asserted unless the relevant
  extended operators, background counterterms, and global-form data are added
  explicitly.
\item Relevant deformations by holomorphic quark masses and by generic
  Higgs-branch expectation values commute with the comparison: after
  integrating out massive fields and matching holomorphic scales, the
  lower-rank theories satisfy the same comparison hypotheses.
\end{enumerate}
\end{hypothesis}

These assumptions are not justified by historical authority.  Their role is
visible in the calculations below.  The magnetic rank and baryon charges are
forced by the baryon map \(Q^{N_c}\leftrightarrow q^{\widetilde N_c}\);
the meson singlet is forced because \(Q\widetilde Q\) is an independent
  electric chiral coordinate; the holomorphic superpotential is the unique
  cubic chiral coupling with the displayed flavor charges and \(R\)-charge
  \(2\), while its canonical coefficient requires the Kähler datum above; the
magnetic gauge anomaly cancellation is a necessary consistency test; and the
local anomaly-polynomial matching below tests the common infrared current
algebra near the identity component of the background group.  These checks do
not construct the continuum infrared fixed point, but they make every
algebraic burden of the duality statement explicit.

\begin{conjecture}[Central conjecture: Seiberg duality for SQCD]
\label{conj:sqcd-seiberg-duality-central}
For four-dimensional \(\mathcal N=1\) SQCD with gauge group \(SU(N_c)\),
\(N_c+2\leq N_f<3N_c\) flavors, zero tree superpotential, and the Wilsonian
coordinate conventions of this chapter, the electric asymptotically free
theory at the origin of the quantum moduli space is conjectured to have the
same infrared QFT as the magnetic \(SU(N_f-N_c)\) theory with fields,
faithful-background action, and superpotential displayed above.  The
conjecture is made under
Hypothesis~\ref{hyp:sqcd-ir-comparison-hypotheses} and is a statement about
the observable sector declared there: local gauge-invariant operators and
correlators, the protected chiral ring after imposing \(F\)-term equations,
holomorphic mass and Higgs deformation families, and partition functions on
admissible backgrounds only after the relevant contact-term convention has
been declared.  It predicts equality of the full anomaly inflow class for the
faithful background data, but the calculation below checks only the local
perturbative anomaly-polynomial component.  In the interval
\(\frac32N_c<N_f<3N_c\) the conjecture asserts a common interacting infrared
SCFT.  In the free magnetic subrange \(N_c+2\leq N_f<\frac32N_c\), with the
lower edge treated separately when \(2N_f=3N_c\), the conjecture asserts the
infrared equivalence after the magnetic free variables, their accidental
symmetries, and the magnetic Wilsonian cutoff have been included in the
infrared datum.  The historical proposal is Seiberg's electric--magnetic
duality \cite{Seiberg1994}; the status in this chapter is central conjecture,
with the following calculations serving as evidence rather than proof.
\end{conjecture}

\begin{conjecture}[Free electric continuation]
\label{conj:sqcd-free-electric-continuation}
For \(N_f\ge 3N_c\) the electric variables define the free electric side of a
separate continuation, rather than another instance of the continuum-pair
statement above.  At \(N_f=3N_c\), the endpoint is treated as a Gaussian edge
because the first nonzero electric beta-function term
\eqref{eq:sqcd-electric-gaussian-edge-beta} makes the small gauge coupling
marginally irrelevant toward the infrared.  The free electric dimensions and
accidental currents are included before any \(\Delta=\frac32R\)
interpretation.  For \(N_f>3N_c\), any electric-side statement must use one of
the ultraviolet choices in
Hypothesis~\ref{hyp:sqcd-ir-comparison-hypotheses}: finite cutoff electric
EFT, specified ultraviolet completion, or electric fields as emergent
infrared coordinates of a magnetic Wilsonian theory with an explicit
nonzero-\(h\) matching cutoff or additional ultraviolet completion.  The
continuation asserts matching only for the protected local sector declared in
that hypothesis and after the limit order stated there.  A UV-complete
electric continuum definition in this range, or a superconformal-dimension
interpretation outside the interacting conformal window, requires additional
input beyond the free electric continuation.
\end{conjecture}

The monograph uses the central conjecture and the free electric continuation
as distinct nonperturbative inputs for the SQCD phase diagram and then exposes
the checks that make each statement sharp: faithful charge assignments, local
anomaly-polynomial matching, deformation matching, and consistency with NSVZ
arithmetic in the range where that arithmetic has fixed-point status.

\paragraph{Magnetic \(R\)-charge and NSVZ consistency.}
With the magnetic charges displayed above, \(U(1)_R\) is free of
\(SU(\widetilde N_c)^2U(1)_R\) gauge anomaly, the magnetic superpotential
has \(R\)-charge \(2\), and the magnetic NSVZ numerator vanishes at the same
candidate superconformal point as the electric numerator.

The \(SU(\widetilde N_c)^2U(1)_R\) anomaly receives \(+\widetilde N_c\)
from the magnetic gaugino.  The two magnetic quark families have fermion
\(R\)-charge
\[
  R(q)-1=R(\widetilde q)-1
  =
  \frac{N_c}{N_f}-1
  =
  -\frac{\widetilde N_c}{N_f}.
\]
Using \(T(\mathbf{fund})=T(\overline{\mathbf{fund}})=\frac12\),
\[
  \widetilde N_c
  +\frac12N_f\left(-\frac{\widetilde N_c}{N_f}\right)
  +\frac12N_f\left(-\frac{\widetilde N_c}{N_f}\right)
  =
  0 .
\]
The superpotential charge is
\[
  R(M)+R(q)+R(\widetilde q)
  =
  2\left(1-\frac{N_c}{N_f}\right)
  +2\frac{N_c}{N_f}
  =
  2 .
\]
At a superconformal fixed point the chapter convention gives
\[
  \gamma_q=3R(q)-2=\frac{3N_c}{N_f}-2 .
\]
The magnetic NSVZ numerator is therefore
\[
  3\widetilde N_c-N_f(1-\gamma_q)
  =
  3(N_f-N_c)
  -N_f\left(3-\frac{3N_c}{N_f}\right)
  =
  0 .
\]
The electric numerator was checked in
Section~\ref{sec:n1-superconformal-sqcd-bookkeeping}.  In the interacting
window \(\frac32N_c<N_f<3N_c\), both descriptions therefore pass the same
fixed-point arithmetic.  Outside that window the same equations remain
algebraic checks on the charge assignments and superpotential, rather than
admission tests for a common interacting fixed point.

\paragraph{Local perturbative anomaly-polynomial matching.}
\label{par:sqcd-seiberg-anomaly-matching}
Normalize cubic nonabelian anomalies so that a left-handed Weyl fermion in
the fundamental of \(SU(N_f)\) contributes \(+1\) and an antifundamental
contributes \(-1\).  Normalize mixed \(SU(N_f)^2U(1)\) anomalies so that a
fundamental or antifundamental contributes the same quadratic index.  The
calculation is performed on liftable product backgrounds for the Lie algebra
of \(G_{FB}^{\rm SQCD}\) and the chosen \(R\)-current.  It checks the free
local part of the anomaly inflow polynomial, not the torsion/global anomaly
classes or contact terms on all quotient bundles.  In this scope, the
electric and magnetic descriptions have the common anomaly coefficients
\[
\begin{array}{c|c}
  \text{anomaly} & \text{coefficient} \\ \hline
  SU(N_f)_L^3 & N_c \\
  SU(N_f)_R^3 & -N_c \\
  SU(N_f)_L^2U(1)_R & -\frac{N_c^2}{N_f} \\
  SU(N_f)_R^2U(1)_R & -\frac{N_c^2}{N_f} \\
  SU(N_f)_L^2U(1)_B & N_c \\
  SU(N_f)_R^2U(1)_B & -N_c \\
  U(1)_B^2U(1)_R & -2N_c^2 \\
  U(1)_R & -N_c^2-1 \\
  U(1)_R^3 & N_c^2-1-\frac{2N_c^4}{N_f^2}.
\end{array}
\]

All traces are over left-handed Weyl fermions.  In the electric theory the
quark and antiquark fermions have \(R\)-charge \(-N_c/N_f\), while the
gaugino has \(R\)-charge \(1\).  Thus, for example,
\[
  SU(N_f)_L^3:\quad N_c,
  \qquad
  SU(N_f)_R^3:\quad -N_c,
\]
and
\[
  SU(N_f)_L^2U(1)_R:
  \quad
  N_c\left(-\frac{N_c}{N_f}\right)
  =
  -\frac{N_c^2}{N_f}.
\]
The right-flavor mixed \(R\)-anomaly is the same, while the baryonic mixed
anomalies are \(N_c\) and \(-N_c\) because \(Q\) and \(\widetilde Q\) have
baryon charges \(+1\) and \(-1\).  The purely abelian traces are
\[
  U(1)_B^2U(1)_R:
  \quad
  2N_cN_f\left(-\frac{N_c}{N_f}\right)
  =
  -2N_c^2,
\]
\[
  \operatorname{Tr}R
  =
  (N_c^2-1)+2N_cN_f\left(-\frac{N_c}{N_f}\right)
  =
  -N_c^2-1,
\]
and
\[
  \operatorname{Tr}R^3
  =
  (N_c^2-1)
  +2N_cN_f\left(-\frac{N_c}{N_f}\right)^3
  =
  N_c^2-1-\frac{2N_c^4}{N_f^2}.
\]

For the magnetic theory set \(t=\widetilde N_c=N_f-N_c\).  The \(q\) and
\(\widetilde q\) fermions have \(R\)-charge \(-t/N_f\), baryon charges
\(\pm N_c/t\), and \(tN_f\) components each.  The mesino in \(M^i{}_j\) has
\[
  R(M)-1
  =
  2\left(1-\frac{N_c}{N_f}\right)-1
  =
  \frac{N_f-2N_c}{N_f}.
\]
For the left cubic flavor anomaly, \(q\) gives \(-t\) and the \(N_f\) right
flavor copies of the mesino give \(+N_f\), hence \(-t+N_f=N_c\).  For the
right cubic anomaly, \(\widetilde q\) gives \(+t\) and the mesino gives
\(-N_f\), hence \(t-N_f=-N_c\).  The left mixed \(R\)-anomaly is
\[
  t\left(-\frac{t}{N_f}\right)
  +
  N_f\left(\frac{N_f-2N_c}{N_f}\right)
  =
  -\frac{(N_f-N_c)^2}{N_f}+N_f-2N_c
  =
  -\frac{N_c^2}{N_f},
\]
and the right mixed \(R\)-anomaly is identical.  The mixed baryonic flavor
anomalies are
\[
  SU(N_f)_L^2U(1)_B:\quad
  t\frac{N_c}{t}=N_c,
  \qquad
  SU(N_f)_R^2U(1)_B:\quad
  -t\frac{N_c}{t}=-N_c .
\]
The \(U(1)_B^2U(1)_R\) trace is
\[
  2tN_f\left(\frac{N_c}{t}\right)^2
  \left(-\frac{t}{N_f}\right)
  =
  -2N_c^2 .
\]
Finally,
\[
  \operatorname{Tr}R
  =
  (t^2-1)+2tN_f\left(-\frac{t}{N_f}\right)
  +N_f^2\left(\frac{N_f-2N_c}{N_f}\right)
  =
  -N_c^2-1,
\]
and
\[
  \operatorname{Tr}R^3
  =
  (t^2-1)
  -\frac{2t^4}{N_f^2}
  +
  \frac{(N_f-2N_c)^3}{N_f}
  =
  N_c^2-1-\frac{2N_c^4}{N_f^2}.
\]
This completes the displayed local anomaly-polynomial matching calculation,
a necessary check on the central conjecture and on the free electric
continuation.  Its validity at \(N_f\ge3N_c\) is algebraic evidence about the
protected current sector, rather than evidence that the electric Lagrangian in
that range has the same ultraviolet status as the asymptotically free
electric theory.  The comparison of all global anomalies on the nonliftable
\(G_{FB}^{\rm SQCD}\)-backgrounds described in
\eqref{eq:sqcd-admissible-background-cocycles} and the choice of
scheme-dependent local background counterterms remain separate compact
background data.

\paragraph{Mass and Higgs deformation tests for SQCD duality.}
Let \(N_f\geq N_c+2\) and set \(\widetilde N_c=N_f-N_c\).  The magnetic
field content and superpotential in
\eqref{eq:sqcd-magnetic-superpotential} pass the following necessary
deformation tests.  This is a bookkeeping calculation inside the proposed
dual description, not a proof of Seiberg duality.

First give the last electric flavor a holomorphic mass
\[
  W_{\mathrm{el,mass}}=m\,M^L{}_L,\qquad L=N_f .
\]
In the magnetic variables the \(F\)-term equation for \(M^L{}_L\) is
\begin{equation}
  \frac1{\mu_*}q_L^\alpha\widetilde q^L_\alpha+m=0 .
  \label{eq:sqcd-mass-deformation-magnetic-fterm}
\end{equation}
For \(\widetilde N_c\geq2\), a rank-one solution Higgses
\[
  SU(\widetilde N_c)\longrightarrow SU(\widetilde N_c-1).
\]
The light magnetic theory has \(N_f-1\) flavors, gauge rank
\[
  \widetilde N_c-1=(N_f-1)-N_c,
\]
and the same meson-quark superpotential form for the light flavor block.
For initial \(\widetilde N_c=2\), the resulting \(SU(1)\) gauge group is
trivial; this is the boundary case leading to the \(N_f=N_c+1\) confining
chiral description rather than to another nonabelian magnetic gauge theory.

Second give the electric theory a rank-\(r\) meson expectation value on the
classical Higgs patch, with \(1\leq r<N_c\):
\[
  M^A{}_B=v_A\delta^A{}_B,\qquad A,B=1,\ldots,r,\qquad v_A\neq0 .
\]
The electric gauge group is Higgsed to \(SU(N_c-r)\) with \(N_f-r\) light
flavors.  In the magnetic variables the same background gives masses
\[
  \frac{v_A}{\mu_*}\,q_A\widetilde q^A,
  \qquad A=1,\ldots,r,
\]
to \(r\) magnetic flavor pairs.  After they are integrated out, the magnetic
gauge rank remains \(\widetilde N_c\), which is precisely the dual rank of
the low-energy electric theory:
\begin{equation}
  (N_f-r)-(N_c-r)=N_f-N_c=\widetilde N_c .
  \label{eq:sqcd-higgs-deformation-rank-test}
\end{equation}

For the mass deformation, split flavor indices as \(i=(a,L)\), with
\(a=1,\ldots,N_f-1\).  The deformed magnetic superpotential is
\[
  W_{\mathrm{mag}}+mM^L{}_L
  =
  \frac1{\mu_*}M^i{}_j q_i^\alpha\widetilde q^j_\alpha
  +mM^L{}_L .
\]
The \(M^L{}_L\) equation is
\eqref{eq:sqcd-mass-deformation-magnetic-fterm}, while the other
\(M\)-equations set the corresponding magnetic quark bilinears to zero on
the light branch.  Choose a local representative
\begin{equation}
  q_L^\alpha=v\,\delta^\alpha{}_{\widetilde N_c},
  \qquad
  \widetilde q^L_\alpha=\widetilde v\,\delta_{\alpha\widetilde N_c},
  \qquad
  v\widetilde v=-\mu_*m ,
  \label{eq:sqcd-magnetic-mass-higgs-representative}
\end{equation}
with the usual \(D\)-flat representative chosen inside the complexified
gauge orbit.  The stabilizer of this nonzero fundamental-antifundamental
pair is \(SU(\widetilde N_c-1)\).  The fields carrying the broken color line
are massive or gauge directions.  The remaining light magnetic quarks have
flavor labels \(a=1,\ldots,N_f-1\) and color labels in
\(SU(\widetilde N_c-1)\), and the surviving singlet block is \(M^a{}_b\).
Restricting the superpotential to these light fields gives
\[
  W_{\mathrm{light}}
  =
  \frac1{\mu_*}M^a{}_b q_a\widetilde q^b ,
\]
which has the same form as \eqref{eq:sqcd-magnetic-superpotential}.  The rank
identity is immediate:
\[
  \widetilde N_c-1=N_f-N_c-1=(N_f-1)-N_c .
\]

For the Higgs deformation, a rank-\(r\) electric meson expectation value may
be represented on the dense Higgs patch by \(r\) independent nonzero
quark-antiquark pairs.  Their common stabilizer in \(SU(N_c)\) is
\(SU(N_c-r)\), and the light electric flavors are the remaining
\(N_f-r\).  In the magnetic superpotential, substituting
\(M^A{}_B=v_A\delta^A{}_B\) gives the displayed mass terms
\((v_A/\mu_*)q_A\widetilde q^A\).  For generic nonzero \(v_A\), these
integrate out exactly \(r\) magnetic flavor pairs.  The magnetic gauge group
is not Higgsed by this operation, so its rank remains
\(\widetilde N_c=N_f-N_c\).  Equation
\eqref{eq:sqcd-higgs-deformation-rank-test} shows that this is the dual rank
of the low-energy electric \(SU(N_c-r)\) theory with \(N_f-r\) flavors.

These deformation tests are necessary consequences of the proposed duality,
not a replacement for the central conjecture.  Their value in this
chapter is that the mass, Higgs, and rank bookkeeping is completely visible:
one can see which fields are eliminated, which gauge group remains, and where
the boundary \(N_f=N_c+1\) description enters.

The \(N_f<N_c\) entries in the phase table below have a different logical
status from the duality entries.  The \(N_f=N_c-1\) ADS term was derived
above from an explicit one-instanton zero-mode and symmetry calculation under
Hypothesis~\ref{hyp:sqcd-one-instanton-calculus}.  The cases
\(N_f<N_c-1\) are then reached by holomorphic decoupling using
the scale matching above and
Proposition~\ref{prop:ads-holomorphic-decoupling-recursion}.  Thus the nonzero
superpotential in the runaway range is not inferred from Seiberg duality; it
is a chiral \(F\)-term result whose direct semiclassical anchor,
finite-dimensional elimination step, and decoupling assumptions are
displayed separately.

\paragraph{SQCD infrared range synthesis.}
\label{par:sqcd-infrared-range-synthesis}
The following table is a synthesis of results and inputs with different
logical status, not a single theorem with a uniform proof.  The
\(N_f<N_c\) range follows from the ADS seed, holomorphic decoupling, and the
displayed finite-dimensional elimination arguments.  The \(N_f=N_c\) and
\(N_f=N_c+1\) entries use the quantum-modified and confining chiral-coordinate
inputs stated earlier in the section.  The electric-asymptotically-free
duality range \(N_c+2\leq N_f<3N_c\) uses the nonperturbative
Seiberg-duality input of
Conjecture~\ref{conj:sqcd-seiberg-duality-central} under
Hypothesis~\ref{hyp:sqcd-ir-comparison-hypotheses}.  The Gaussian edge and
the strict free electric continuation use the separate status of
Conjecture~\ref{conj:sqcd-free-electric-continuation}.  The one-loop signs
below are Wilsonian coordinate checks on the proposed infrared variables; by
themselves they supply necessary algebraic evidence rather than continuum
infrared fixed points or electric-magnetic equivalence.
\begin{description}
\item[\(N_f<N_c\).]
Status: ADS/holomorphic-decoupling result under the instanton and chiral
coordinate hypotheses already stated.  The ADS superpotential is generated.
With no tree-level mass terms there is no supersymmetric vacuum at finite
meson expectation value; the potential has a runaway direction on the
maximal-rank meson patch.
\item[\(N_f=N_c\).]
Status: quantum-modified chiral-coordinate input checked against the
decoupling from the \(N_f=N_c+1\) description.  There is no generated
superpotential, but the chiral coordinate ring is quantum modified:
\[
  \det M-B\widetilde B=\Lambda_h^{2N_c}.
\]
\item[\(N_f=N_c+1\).]
Status: confining chiral-coordinate input with the mass-decoupling and anomaly
checks displayed above.  The infrared chiral description is in terms of mesons
and baryons.  If \(B_i\) and \(\widetilde B^j\) denote the baryons with one
flavor index left over, the exact superpotential is
\begin{equation}
  W_{\mathrm{conf}}
  =
  \frac{1}{\Lambda_h^{2N_c-1}}
  \left(B_iM^i{}_j\widetilde B^j-\det M\right).
  \label{eq:sqcd-sconfining-superpotential}
\end{equation}
\item[\(N_c+2\leq N_f<\frac32N_c\).]
Status: Seiberg-duality input plus perturbative sign check in the magnetic
variables.  The magnetic gauge group \(SU(N_f-N_c)\) is infrared free, since
its one-loop coefficient is
\[
  b_{0,\mathrm{mag}}
  =
  3(N_f-N_c)-N_f
  =
  2N_f-3N_c<0 .
\]
The appropriate low-energy variables are the magnetic quarks, magnetic
gluons, and meson singlets.  The magnetic Lagrangian is used as the
infrared-free Wilsonian description in the declared low-energy window, with
the free-field accidental currents included in the infrared datum.
\item[\(2N_f=3N_c\).]
When this equality is compatible with integrality, the entry is the lower
endpoint of the interacting-window bookkeeping and the upper endpoint of the
free magnetic sign check.  The magnetic one-loop coefficient vanishes, the
electric meson saturates the chiral unitarity bound, and the superconformal
\(R\)-charge arithmetic alone no longer identifies an interacting fixed point
without adding the accidental free-field current.
\item[\(\frac32N_c<N_f<3N_c\).]
Status: Seiberg-duality input plus SCFT-current and accidental-symmetry
hypotheses.  Both electric and magnetic variables are asymptotically free and
flow, under that input, to the same interacting nonabelian Coulomb-phase SCFT.
The lower endpoint is detected by the electric meson unitarity bound, and the
upper endpoint by the electric loss of asymptotic freedom.
\item[\(N_f=3N_c\).]
Status: Gaussian electric edge of the asymptotically free family and boundary
case of the free electric continuation.  The electric one-loop coefficient
vanishes, but the first nonzero electric beta-function term
\eqref{eq:sqcd-electric-gaussian-edge-beta} is positive in the chapter's
\(\beta_g=\mu\,dg/d\mu\) convention, so the small electric coupling is
marginally irrelevant toward the infrared.  The anomaly comparison
\eqref{eq:sqcd-free-field-a-comparison} has zero difference from the free
electric value, and fixed-point dimension formulae are interpreted with the
free electric variables and their accidental currents rather than as an
interacting SCFT claim.  The magnetic gauge coupling is asymptotically free
at this endpoint, but the canonical magnetic superpotential has no nonzero
weak-coupling gauge--Yukawa asymptotically-free ray by
\eqref{eq:sqcd-magnetic-yukawa-ratio-flow}; it is not a replacement
derivation of the electric Gaussian edge.
\item[\(N_f>3N_c\).]
Status: strict free electric continuation.  The electric gauge coupling is
infrared free in the perturbative Wilsonian sense, while the electric
Lagrangian by itself is a finite cutoff EFT, a UV-completed theory, or an
emergent infrared coordinate system supplied by a magnetic Wilsonian
definition with a declared matching cutoff and Yukawa value.  Magnetic gauge
asymptotic freedom is not enough: for nonzero canonical \(h\), the one-loop
system \eqref{eq:sqcd-magnetic-yukawa-beta-one-loop}--
\eqref{eq:sqcd-magnetic-yukawa-ratio-flow} has a Yukawa UV obstruction rather
than a positive nonzero asymptotically-free ray.  The phase entry therefore
asserts only the protected low-energy sector specified in
Hypothesis~\ref{hyp:sqcd-ir-comparison-hypotheses}, with the order of
thermodynamic, infrared, and cutoff limits stated there.  The same rank,
charge, anomaly, NSVZ, and deformation arithmetic remains a consistency
ledger, while the continuum-pair statement of
Conjecture~\ref{conj:sqcd-seiberg-duality-central} requires the separate
range and ultraviolet hypotheses stated above.
\end{description}

\section{Klebanov--Witten Conifold SCFT}
\label{sec:n1-klebanov-witten-conifold}

Let \(N\geq2\).  The Klebanov--Witten theory has gauge group
\[
  G_{\mathrm{KW}}=SU(N)_1\times SU(N)_2
\]
and four chiral superfields
\[
  A_i\in(\mathbf N,\overline{\mathbf N}),\qquad
  B_j\in(\overline{\mathbf N},\mathbf N),
  \qquad i,j=1,2 .
\]
Equivalently, one may start from the unitary
\(U(N)_1\times U(N)_2\) presentation.  Its diagonal \(U(1)\) is a free
center-of-mass vector multiplet, and the interacting nonabelian sector is
the \(SU(N)_1\times SU(N)_2\) theory displayed above.  The relative abelian
factor is useful for describing the rank-one abelian quotient below.

The index \(i\) is a doublet index for a global \(SU(2)_A\), and \(j\) is a
doublet index for a global \(SU(2)_B\).  The baryonic symmetry \(U(1)_B\) is
normalized here by
\[
  q_B(A_i)=+1,\qquad q_B(B_j)=-1 .
\]
The tree superpotential is
\begin{equation}
  W_{\mathrm{KW}}
  =
  h\,\epsilon^{ij}\epsilon^{kl}
  \operatorname{tr}(A_iB_kA_jB_l),
  \label{eq:kw-superpotential}
\end{equation}
where the trace is over the gauge indices in the alternating product
\[
  \mathbf N_1\xrightarrow{A_i}\mathbf N_2
  \xrightarrow{B_k}\mathbf N_1
  \xrightarrow{A_j}\mathbf N_2
  \xrightarrow{B_l}\mathbf N_1 .
\]
The manifest continuous global symmetry at the superpotential point is
\[
  SU(2)_A\times SU(2)_B\times U(1)_B\times U(1)_R .
\]
There is also a charge-conjugation symmetry exchanging the two gauge factors
and the \(A\)- and \(B\)-multiplets when the two gauge couplings are
exchanged.

\begin{hypothesis}[KW Lagrangian and branch assumptions]
\label{hyp:kw-lagrangian-branch-assumptions}
The algebraic calculations in this section use the four-dimensional
\(\mathcal N=1\) Lagrangian with gauge group, matter representations, trace
normalization, and superpotential displayed above.  Statements about the
mesonic vacuum branch use the classical \(F\)- and \(D\)-flat quotient of
that Lagrangian and then restrict to the commuting branch where the
bifundamental matrices can be simultaneously reduced to diagonal
representatives.  Statements about protected dimensions use
Hypothesis~\ref{hyp:n1-scft-bookkeeping-input}.  Claims about the full
continuum conformal manifold are isolated in
Hypothesis~\ref{hyp:kw-exact-marginality-input} and in the
exact-marginality rank calculation below.
\end{hypothesis}

\paragraph{KW \(F\)-term equations.}
The \(F\)-term equations of \(W_{\mathrm{KW}}\) are
\begin{equation}
  A_1B_jA_2=A_2B_jA_1,\qquad j=1,2,
  \label{eq:kw-f-term-a}
\end{equation}
and
\begin{equation}
  B_1A_iB_2=B_2A_iB_1,\qquad i=1,2 .
  \label{eq:kw-f-term-b}
\end{equation}

Expanding the antisymmetric tensors and using cyclicity of the trace gives
\[
  W_{\mathrm{KW}}
  =
  2h\,\operatorname{tr}(A_1B_1A_2B_2-A_1B_2A_2B_1).
\]
Varying with respect to \(A_1\) and moving \(\delta A_1\) to the left inside
the trace gives
\[
  \delta_{A_1}W_{\mathrm{KW}}
  =
  2h\,\operatorname{tr}\bigl[
    \delta A_1(B_1A_2B_2-B_2A_2B_1)
  \bigr],
\]
so \(B_1A_2B_2=B_2A_2B_1\).  The variation with respect to \(A_2\) gives the
same equation with \(A_1\) and \(A_2\) exchanged.  Varying with respect to
\(B_1\) and \(B_2\) in the same way gives
\[
  A_2B_2A_1=A_1B_2A_2,\qquad
  A_1B_1A_2=A_2B_1A_1,
\]
which is exactly \eqref{eq:kw-f-term-a}.  These are matrix equations in the
appropriate bifundamental Hom-spaces.

\paragraph{KW anomaly and superpotential \(R\)-charge.}
Assume the \(SU(2)_A\times SU(2)_B\) symmetry and the exchange symmetry, so
that
\[
  R(A_i)=R(B_j)=r .
\]
The simultaneous conditions that \(U(1)_R\) have no \(SU(N)_a^2U(1)_R\)
anomaly and that \(W_{\mathrm{KW}}\) have \(R\)-charge \(2\) give
\[
  r=\frac12 .
\]
For either gauge factor the gaugino contributes \(T(\operatorname{adj})=N\)
to the \(SU(N)^2U(1)_R\) anomaly coefficient.  Each of the two \(A\)-fields
and each of the two \(B\)-fields is a fundamental or antifundamental for that
gauge factor and has \(N\) copies from the other gauge index.  With
\(T(\mathbf N)=T(\overline{\mathbf N})=\frac12\), the matter contribution is
\[
  4\cdot \frac{N}{2}(r-1)=2N(r-1).
\]
The anomaly coefficient is therefore
\[
  N+2N(r-1)=N(2r-1),
\]
which vanishes exactly at \(r=\frac12\).  The superpotential contains two
\(A\)-fields and two \(B\)-fields, so its \(R\)-charge is \(4r\), also equal
to \(2\) exactly at \(r=\frac12\).

\begin{remark}[KW NSVZ numerator]
\label{rem:kw-nsvz-numerator}
At \(r=\frac12\), as fixed by the anomaly and superpotential-charge
calculation above, the
chiral-multiplet anomalous dimensions in the superconformal convention are
\[
  \gamma_A=\gamma_B=-\frac12 ,
\]
and the NSVZ numerator of each gauge factor vanishes.
Equation~\eqref{eq:n1-gamma-from-r} gives
\[
  \gamma_A=\gamma_B=3\cdot\frac12-2=-\frac12 .
\]
For either \(SU(N)\) factor, the total matter index is
\[
  4\cdot N\cdot\frac12=2N.
\]
Hence the NSVZ numerator is
\[
  3N-2N(1-\gamma_A)
  =
  3N-2N\left(1+\frac12\right)
  =
  0 .
\]
The scalar dimension of each \(A_i\) and \(B_j\) is therefore
\[
  \Delta(A_i)=\Delta(B_j)=1+\frac12\gamma_A=\frac34
  =
  \frac32R(A_i).
\]
\end{remark}

\paragraph{KW beta-function rank count.}
Work in the \(SU(2)_A\times SU(2)_B\times U(1)_B\)-preserving source-coordinate system
near the Klebanov--Witten point, and assume the NSVZ coordinate convention of
this chapter with finite denominators for both gauge couplings.  Let
\(\gamma_A\) and \(\gamma_B\) be the anomalous dimensions of the \(A\)- and
\(B\)-doublets in the monograph convention.  Then the two gauge beta-function
numerators and the quartic-superpotential marginality defect are all
proportional to the same expression:
\begin{equation}
  \mathcal B_1=\mathcal B_2=N(1+\gamma_A+\gamma_B),
  \qquad
  \mathcal B_h=1+\gamma_A+\gamma_B.
  \label{eq:kw-beta-rank-common-defect}
\end{equation}
Consequently the three local vanishing conditions have rank one in this
symmetry-preserving chart.  At the exchange-symmetric point
\(\gamma_A=\gamma_B=\gamma\), the common condition is
\(\gamma=-\frac12\).

For the first gauge group, each of the two \(A\)-fields is a fundamental with
\(N\) copies from the second gauge index, and each of the two \(B\)-fields is
an antifundamental with the same multiplicity.  Hence
\[
  \sum_iT_1(R_i)(1-\gamma_i)
  =
  2\cdot\frac N2(1-\gamma_A)
  +
  2\cdot\frac N2(1-\gamma_B)
  =
  N(2-\gamma_A-\gamma_B).
\]
The NSVZ numerator of the first gauge factor is therefore
\[
  \mathcal B_1
  =
  3N-N(2-\gamma_A-\gamma_B)
  =
  N(1+\gamma_A+\gamma_B).
\]
The same count holds for the second gauge factor, with fundamentals and
antifundamentals interchanged, so \(\mathcal B_2=\mathcal B_1\).

For the superpotential term, nonrenormalization means that the local
condition for a quartic chiral coordinate to be marginal is the dimension
condition on the chiral operator.  The operator
\(\operatorname{tr}(A_iB_kA_jB_l)\) has dimension
\[
  2\left(1+\frac{\gamma_A}{2}\right)
  +
  2\left(1+\frac{\gamma_B}{2}\right)
  =
  4+\gamma_A+\gamma_B .
\]
Since a four-dimensional superpotential integrand must have dimension \(3\),
the marginality defect is
\[
  \mathcal B_h
  =
  (4+\gamma_A+\gamma_B)-3
  =
  1+\gamma_A+\gamma_B .
\]
Equivalently, the beta function of a dimensionless quartic coupling is
proportional to this defect, up to the sign convention for whether the flow
parameter is an energy scale or a length scale.  Only the zero locus is used
here.

Thus \(\mathcal B_1\), \(\mathcal B_2\), and \(\mathcal B_h\) impose a single
scalar condition in the symmetric chart.  This is a necessary local
beta-function count.  Existence of the continuum SCFT, realization of the
common zero by an RG trajectory, and nonsingularity of the local zero locus are
the additional inputs isolated in
Hypothesis~\ref{hyp:kw-exact-marginality-input}.

The anomaly and NSVZ calculations fix the symmetric point.  The trial-charge
check uses Quoted Theorem~\ref{qt:n1-a-maximization-boundary} and is
conditional on completeness of the trial-current basis for the preserved
symmetry.  In the sector preserving
\(SU(2)_A\times SU(2)_B\), mixing with Cartans of the two non-Abelian flavor
groups is not allowed because it would choose directions in the doublets.  A
common axial shift of \(R(A)\) and \(R(B)\) violates the gauge-anomaly and
quartic-superpotential constraints.  Thus the only Abelian mixing direction in
this sector is the baryonic current, normalized by \(q_B(A)=+1\),
\(q_B(B)=-1\):
\[
  R_s(A_i)=\frac12+s,\qquad
  R_s(B_j)=\frac12-s .
\]
The gauge-anomaly and superpotential constraints impose only
\(R_s(A)+R_s(B)=1\).  The trial central charge is computed from Weyl fermion
\(R\)-charges.  The two gaugino adjoints have fermion \(R\)-charge \(1\);
the two \(A\)-fermions have charge \(-\frac12+s\); the two \(B\)-fermions
have charge \(-\frac12-s\).  Therefore
\[
  \operatorname{Tr}R_s=-2,
\]
and
\[
  \operatorname{Tr}R_s^3
  =
  2(N^2-1)
  +
  2N^2\left[
    \left(-\frac12+s\right)^3
    +
    \left(-\frac12-s\right)^3
  \right]
  =
  \frac32N^2-2-6N^2s^2 .
\]
Thus
\[
  a(s)
  =
  \frac{3}{32}
  \left(3\operatorname{Tr}R_s^3-\operatorname{Tr}R_s\right)
  =
  \frac{3}{32}
  \left(\frac92N^2-4-18N^2s^2\right),
\]
which has stationary point \(s=0\) and Hessian
\[
  a''(0)=-\frac{27}{8}N^2 .
\]
The current two-point relation
\eqref{eq:n1-a-max-hessian-current-two-point} gives the same sign:
\[
  \operatorname{Tr}R_\ast B^2=-2N^2,\qquad
  \tau_{BB}=6N^2,\qquad
  -\frac9{16}\tau_{BB}=-\frac{27}{8}N^2 .
\]
Assuming no accidental current is present in this sector, the
\(a\)-maximization theorem therefore selects \(s=0\).  Under these hypotheses
the superconformal \(R\)-charge is the symmetric one:
\[
  R(A_i)=R(B_j)=\frac12 .
\]

\begin{remark}[KW central charges]
\label{rem:kw-central-charges}
At the symmetric superconformal \(R\)-charge,
\[
  a_{\mathrm{KW}}=\frac{27}{64}N^2-\frac38,\qquad
  c_{\mathrm{KW}}=\frac{27}{64}N^2-\frac14 .
\]
For a four-dimensional \(\mathcal N=1\) SCFT,
\[
  a=\frac{3}{32}(3\operatorname{Tr}R_f^3-\operatorname{Tr}R_f),
  \qquad
  c=\frac{1}{32}(9\operatorname{Tr}R_f^3-5\operatorname{Tr}R_f),
\]
where the traces run over Weyl fermions.  The fermion trace data at \(s=0\)
are
\[
  \operatorname{Tr}R_f
  =
  2(N^2-1)+4N^2\left(-\frac12\right)
  =
  -2
\]
and
\[
  \operatorname{Tr}R_f^3
  =
  2(N^2-1)+4N^2\left(-\frac18\right)
  =
  \frac32N^2-2 .
\]
Substitution gives the displayed values.
\end{remark}

The basic mesonic chiral operators are
\[
  W_{ij}=A_iB_j .
\]
For a single rank-one abelian sector in the \(U(1)\times U(1)\)
presentation, the diagonal \(U(1)\) decouples and the complexified
anti-diagonal gauge group acts with charges \(+1\) on the \(A_i\) and
\(-1\) on the \(B_j\).  The four coordinates \(W_{ij}\) are invariant and
obey one quadratic relation.

For the rank-one abelian quotient just described, the mesonic coordinate
ring is
\[
  \C[W_{11},W_{12},W_{21},W_{22}]
  /
  (W_{11}W_{22}-W_{12}W_{21}) .
\]
Write \(W_{ij}=A_iB_j\) with commuting complex variables
\(A_1,A_2,B_1,B_2\).  Then
\[
  W_{11}W_{22}-W_{12}W_{21}
  =
  A_1B_1A_2B_2-A_1B_2A_2B_1
  =
  0 .
\]
Conversely, the invariant functions of a \(\C^\times\) action with charges
\(+1,+1,-1,-1\) are generated by the degree-zero monomials \(A_iB_j\).  The
only rank-one relation among the entries of the \(2\times2\) matrix
\((W_{ij})\) is its determinant.  Hence the quotient is the affine conifold
\[
  \det(W)=0 .
\]

On the mesonic commuting branch for general \(N\), the matrices \(A_iB_j\)
can be diagonalized simultaneously at generic points, and the Weyl group
permutes the \(N\) eigenvalue quadruples.  The corresponding branch is
\[
  \operatorname{Sym}^N\{W_{11}W_{22}-W_{12}W_{21}=0\}.
\]
This statement concerns the commuting mesonic branch; baryonic branches and
noncommuting chiral-ring strata require additional invariant-theory data.

Under Hypothesis~\ref{hyp:n1-scft-bookkeeping-input} and the
\(a\)-maximization boundary
Quoted Theorem~\ref{qt:n1-a-maximization-boundary}, the protected operator
dimensions follow from the selected superconformal \(R\)-charges.  The mesons
\(W_{ij}\) transform as
\((\mathbf2,\mathbf2)\) under \(SU(2)_A\times SU(2)_B\) and have
\[
  R(W_{ij})=1,\qquad \Delta(W_{ij})=\frac32 .
\]
The baryons
\[
  \mathcal B_A=\det A_i,\qquad
  \mathcal B_B=\det B_j
\]
have
\[
  R(\mathcal B_A)=R(\mathcal B_B)=\frac N2,\qquad
  \Delta(\mathcal B_A)=\Delta(\mathcal B_B)=\frac{3N}{4},
\]
and baryonic charges \(+N\) and \(-N\) in the normalization above.
Symmetric products
\[
  \operatorname{tr}(A_{i_1}B_{j_1}\cdots A_{i_\ell}B_{j_\ell})
\]
have \(R=\ell\) and \(\Delta=\frac32\ell\) when they represent nonzero
chiral-primary classes.

\begin{hypothesis}[Local \(\mathcal N=1\) exact-marginality chart]
\label{hyp:n1-local-exact-marginality-chart}
Let \(\mathcal T_0\) be a four-dimensional \(\mathcal N=1\) SCFT with a
chosen global-symmetry subalgebra \(\mathfrak f\) to be preserved.  A local
holomorphic marginal-source-coordinate system consists of complex coordinates
\[
  u=(u^1,\ldots,u^m)
\]
for gauge and superpotential sources coupled to \(\mathfrak f\)-invariant
chiral operators of superpotential dimension \(3\), together with a
holomorphic beta map
\[
  \mathcal B:U\subset\mathbb C^m\longrightarrow\mathbb C^r
\]
whose zero set is exactly the set of conformal theories in this source
chart.  Redundant directions from holomorphic field-coordinate changes have
either been quotient out or fixed by a declared source normalization.  The
map \(\mathcal B\) includes the gauge beta functions, superpotential beta
functions, and any current-moment-map obstruction for a broken conserved
current.  If \(d\mathcal B\) has constant rank \(r\) along the zero set, the
local conformal manifold in this chart is the complex manifold
\(\mathcal B^{-1}(0)\) of dimension \(m-r\).
\end{hypothesis}

Under this hypothesis, the dimension statement is the holomorphic
implicit-function theorem applied to the finite-dimensional source-coordinate system.
The QFT content is not the theorem of complex analysis; it is the hypothesis
that the beta map and the quotient by redundant coordinates faithfully
describe the continuum SCFT family.

\begin{hypothesis}[KW exact-marginality input]
\label{hyp:kw-exact-marginality-input}
Near the Klebanov--Witten point, use local holomorphic sources
\[
  u=(\tau_1,\tau_2,h)
\]
for the two \(SU(N)\) gauge couplings and the quartic superpotential in
\eqref{eq:kw-superpotential}, restricted to the
\(SU(2)_A\times SU(2)_B\times U(1)_B\)-preserving chart.  Assume the
continuum SCFT exists in this neighborhood, the NSVZ Wilsonian coordinate
analysis of this chapter computes the vanishing conditions for these sources,
and no additional exactly marginal obstruction is present after the
background contact-term and source-normalization data are fixed.  Finally
assume the common defect
\[
  E(u)=1+\gamma_A(u)+\gamma_B(u)
\]
has nonzero differential at the symmetric point in the chosen source-coordinate system.
\end{hypothesis}

\paragraph{KW local conformal manifold from rank one.}
Assume Hypothesis~\ref{hyp:kw-exact-marginality-input}.  In the
\(SU(2)_A\times SU(2)_B\times U(1)_B\)-preserving source-coordinate system, the
Klebanov--Witten conformal locus is locally a complex surface in the
source space with coordinates \((\tau_1,\tau_2,h)\).

The rank-count calculation above gives, in this chart,
\[
  \mathcal B_1=N E,\qquad
  \mathcal B_2=N E,\qquad
  \mathcal B_h=E,
  \qquad
  E=1+\gamma_A+\gamma_B .
\]
Thus the zero set of the three beta functions is the single equation
\[
  E(u)=0 .
\]
By Hypothesis~\ref{hyp:kw-exact-marginality-input}, \(dE\neq0\) at the
Klebanov--Witten point.  The holomorphic implicit-function theorem therefore
identifies the local zero set with a smooth complex hypersurface in the
three-dimensional source-coordinate system.  Its local complex dimension is
\[
  3-1=2 .
\]
The proof uses no holographic construction.  Its nontrivial QFT assumptions
are exactly those isolated in
Hypothesis~\ref{hyp:kw-exact-marginality-input}: existence of the continuum
SCFT chart, completeness of the beta-function analysis, and absence of hidden
current or contact-term obstructions in the preserved-symmetry sector.

The result is a conditional field-theoretic reconstruction of the
Leigh--Strassler rank count for this example.  The chapter has proved the
anomaly, NSVZ, chiral-ring, and rank-one beta-function arithmetic, and it has
applied the \(a\)-maximization theorem under the stated current-basis and
no-accidental-symmetry hypotheses.  It has not constructed the
nonperturbative continuum SCFT from a regulator.

\section{Klebanov--Strassler Cascade}
\label{sec:n1-klebanov-strassler-cascade}

Fix integers \(M>0\) and \(N>M\).  The ultraviolet theory used in the
Klebanov--Strassler cascade has gauge group
\[
  SU(N+M)_1\times SU(N)_2
\]
with the same bifundamental matter and quartic superpotential as
\eqref{eq:kw-superpotential}, now with rectangular gauge indices.  Treating
the anomalous dimensions of the four bifundamentals as equal to a common
\(\gamma\), the NSVZ numerators are
\begin{equation}
  \mathcal B_1
  =
  3(N+M)-2N(1-\gamma),
  \qquad
  \mathcal B_2
  =
  3N-2(N+M)(1-\gamma).
  \label{eq:ks-nsvz-numerators}
\end{equation}
At the Klebanov--Witten value \(\gamma=-\frac12\),
\[
  \mathcal B_1=3M,\qquad
  \mathcal B_2=-3M .
\]
Since the NSVZ beta function has the sign
\(\beta_g=-g^3\mathcal B/(16\pi^2(1-\cdots))\), the larger gauge factor is
driven to stronger coupling toward the infrared in this local coordinate,
while the smaller factor is driven in the opposite direction.  The cascade
idealizes the strongly coupled step by applying Seiberg duality to the
larger factor.

\begin{hypothesis}[Cascade input assumptions]
\label{hyp:ks-cascade-input-assumptions}
The cascade discussion assumes the same supersymmetric Wilsonian scheme and
NSVZ coordinate convention used earlier in the chapter.  The displayed
numerator signs are perturbative-coordinate diagnostics near the
Klebanov--Witten anomalous dimension, not a proof that the flow follows a
unique trajectory.  Each rank-reducing step additionally assumes Seiberg
duality for the strongly coupled node with the other node treated during the
matching as a weakly gauged flavor symmetry, and assumes that integrating out
the mesons in the magnetic description is legitimate in the Wilsonian
neighborhood being used.  These duality assumptions are the SQCD central
conjecture of Conjecture~\ref{conj:sqcd-seiberg-duality-central}, applied
with the neighboring gauge factor included in the declared background data and
with the dualized node inside the stated \(N_f\ge N_c+2\) range.  The
near-end \(N_f=N_c+1\) case uses the separately stated confining
chiral-coordinate input rather than the nonabelian magnetic duality
conjecture.
\end{hypothesis}

\paragraph{Cascade rank step.}
Assume the Seiberg-duality transformation for the first gauge factor while
the second gauge factor is treated as a weakly gauged flavor symmetry during
the dualization, and suppose first that \(N-M\ge2\).  Then
\[
  SU(N+M)\times SU(N)
  \quad\longmapsto\quad
  SU(N)\times SU(N-M),
\]
up to relabeling of the two factors and matching of the quartic
superpotential.

The first gauge factor \(SU(N+M)\) sees two \(A\)-fields and two \(B\)-fields.
Because each carries an \(SU(N)\) index of the second gauge group, it has
\[
  N_f=2N
\]
flavors in the SQCD sense.  Seiberg duality maps an electric
\(SU(N_c)\) theory with \(N_f\) flavors to a magnetic
\(SU(N_f-N_c)\) theory.  Here
\[
  N_f-N_c=2N-(N+M)=N-M .
\]
The untouched second gauge factor is \(SU(N)\).  Relabeling the larger
factor first gives \(SU(N)\times SU(N-M)\).  The electric mesons of the
dualized node are composites
\[
  M_{ij}=A_iB_j .
\]
The magnetic superpotential has the local form
\begin{equation}
  W_{\mathrm{mag}}
  =
  h\,\epsilon^{ij}\epsilon^{kl}\operatorname{tr}(M_{ik}M_{jl})
  +
  \frac1{\mu_*}\operatorname{tr}(M_{ij}a_i b_j),
  \label{eq:ks-magnetic-superpotential-local}
\end{equation}
where \(a_i,b_j\) are magnetic bifundamentals and \(\mu_*\) is the matching
scale used to assign dimension one to \(M_{ij}\).  The quadratic form
\[
  (M_{ij})\longmapsto
  \epsilon^{ij}\epsilon^{kl}\operatorname{tr}(M_{ik}M_{jl})
\]
is nondegenerate on the four mesons.  Thus the \(F\)-term equations of
\eqref{eq:ks-magnetic-superpotential-local} solve \(M_{ij}\) linearly in
the bilinears \(a_i b_j\).  Substitution gives again a quartic
superpotential proportional to
\[
  \epsilon^{ij}\epsilon^{kl}\operatorname{tr}(a_i b_k a_j b_l),
\]
with a coefficient depending on the normalization of \(M_{ij}\) and
\(\mu_*\).

Thus the rank difference remains \(M\), while the common rank parameter
drops by \(M\).  If
\[
  N=qM+r,\qquad 0\leq r<M ,
\]
then after \(q\) formal cascade steps the pair has reached
\[
  SU(M+r)\times SU(r),
\]
where the second factor is absent when \(r=0\).  The integer \(q\) is the
number of rank-reducing duality steps, and \(r=N\bmod M\) is the residual
common rank parameter.

\paragraph{KS \(R\)-anomaly remnant.}
For the unequal-rank theory with the field charges
\[
  R(A_i)=R(B_j)=\frac12,\qquad R(\lambda_1)=R(\lambda_2)=1,
\]
the continuous \(U(1)_R\) has gauge-anomaly coefficients
\[
  \mathcal A_1=M,\qquad \mathcal A_2=-M
\]
for \(SU(N+M)_1^2U(1)_R\) and \(SU(N)_2^2U(1)_R\).  Consequently only the
\(\mathbb Z_{2M}\) subgroup is an honest symmetry with fixed theta angles.

For \(SU(N+M)_1\), the gaugino contributes \(N+M\).  Each chiral fermion in
an \(A_i\) or \(B_j\) multiplet has \(R\)-charge \(-\frac12\), and the total
matter index seen by the first gauge group is
\[
  4\cdot N\cdot\frac12=2N .
\]
Therefore
\[
  \mathcal A_1=(N+M)-\frac12(2N)=M .
\]
For \(SU(N)_2\), the total matter index is instead
\[
  4\cdot (N+M)\cdot\frac12=2(N+M),
\]
so
\[
  \mathcal A_2=N-\frac12\bigl(2(N+M)\bigr)=-M .
\]
In the normalization where pure \(SU(K)\) supersymmetric Yang--Mills leaves
\(\mathbb Z_{2K}\), an \(R\)-rotation by angle \(\varphi\) shifts the two
theta angles by
\[
  \theta_1\mapsto\theta_1+2M\varphi,\qquad
  \theta_2\mapsto\theta_2-2M\varphi .
\]
Because each theta angle has period \(2\pi\), the rotation is a symmetry at
fixed theta angles exactly when
\[
  2M\varphi\in2\pi\mathbb Z .
\]
Modulo the \(2\pi\)-periodicity of the \(R\)-rotation parameter, this leaves
\(2M\) elements, namely \(\mathbb Z_{2M}\).

\begin{conjecture}[Cascade use of the SQCD central conjecture]
\label{conj:ks-seiberg-duality-cascade}
The cascade rank step describes the infrared chiral data, local
anomaly-polynomial data, and holomorphic deformations of the successive
Wilsonian theories provided Conjecture~\ref{conj:sqcd-seiberg-duality-central}
applies to the strongly coupled node with the neighboring gauge group treated
as a weakly gauged flavor symmetry during the matching.  The matching includes
the faithful flavor-baryon background data inherited from the node, the
declared \(R\)-current background convention, holomorphic scale coordinates,
and the quartic superpotential after the massive mesons are eliminated.  It
requires the node to lie in the SQCD range where that central conjecture is
available; the \(N_f=N_c+1\) endpoint is instead governed by the confining
chiral-coordinate input.  It does not assert equality of all line/defect
sectors or all quotient-background torsion anomalies unless those data are
added to the SQCD conjecture.
\end{conjecture}

This is a conditional use of the SQCD central conjecture, not an independent
derivation.  The monograph keeps the visible checks: the magnetic rank is
\(N_f-N_c\), the rank difference is preserved, the alternating quartic
superpotential is regenerated, and the local anomaly-polynomial assignment is
unchanged after the duality map.

For \(r=0\), the cascade ends at pure \(SU(M)\) supersymmetric Yang--Mills.
The anomalous chiral symmetry leaves a \(\mathbb Z_{2M}\) gaugino phase
symmetry, and gaugino condensation breaks it to \(\mathbb Z_2\), producing
\(M\) massive vacua in the finite-volume/gap sense used earlier in this
chapter.  For \(0<r<M\), the final strongly coupled theory contains an ADS
or quantum-modified chiral coordinate, depending on the precise endpoint
rank.  This is only a field-theoretic endpoint summary in the present
monograph; any geometric interpretation is outside the argument and is not
used as evidence for the QFT dynamics.

\section{Status Boundary}

The arguments in this chapter determine holomorphic coordinates from symmetry,
anomaly, decoupling, and semiclassical input.  Turning them into a theorem for
a four-dimensional continuum QFT requires a nonperturbative construction of
the regulated supersymmetric gauge theory, a proof of the required mass gaps
or moduli-space descriptions, and a Wilsonian scheme in which the chiral
coordinates above are defined as continuum limits.

\begin{openproblem}[Nonperturbative \(\mathcal N=1\) Wilsonian scheme]
\label{op:n1-nonperturbative-wilsonian-scheme}
Construct a four-dimensional regulator and continuum limit for
\(\mathcal N=1\) gauge theories that preserves enough supersymmetric and gauge
Ward identities to make the holomorphic Wilsonian coordinate arguments above
proved statements.
\end{openproblem}
