\chapter{Supersymmetry Algebras and Representation Data}
\label{ch:susy-algebras-representations}

\ConventionsLorentzian

Supersymmetry is an extension of spacetime symmetry by odd generators.  The
word ``odd'' refers to a \(\mathbb Z_2\)-graded algebra: even generators
commute by ordinary commutators, odd generators anticommute with odd
generators, and mixed brackets are ordinary commutators.  This chapter fixes
the representation-theoretic data before introducing superfields or
Lagrangians.

\section{Particle Multiplets and Field-Variable Multiplets}

\begin{definition}[Particle supersymmetry representation]
\label{def:susy-particle-representation}
A particle representation of supersymmetry is a positive-energy unitary
representation of the real super-Poincare algebra on a \(\mathbb Z_2\)-graded
Hilbert space, or on a one-particle spectral subspace of such a Hilbert
space.  The odd generators are closed operators whose anticommutators are
understood on a common invariant domain.  A particle supermultiplet is an
irreducible representation after the mass orbit, spin or helicity data,
internal charges, and central charges have been fixed.
\end{definition}

\begin{definition}[Field-variable supersymmetry representation]
\label{def:susy-field-variable-representation}
A field-variable representation of supersymmetry is an action of the
super-Poincare algebra, or of its supertranslation subalgebra, by derivations
on a space of classical field variables, superfields, antifields, or local
functionals.  It is off shell when the algebra closes without imposing
Euler--Lagrange equations.  It is on shell when closure holds only modulo
equations of motion, gauge transformations, or both.  Auxiliary fields are
field variables added to obtain off-shell closure; they are not particle
states by that fact alone.
\end{definition}

The two notions of multiplet have different ambient categories.  Particle
multiplets live in Hilbert-space representation theory and are constrained by
positivity of the supercharge anticommutator.  Off-shell field multiplets live
in a coordinate description of local dynamics.  A relation between them is a
derived statement: one must choose an action or equations of motion, impose
constraints, quotient gauge redundancy if present, quantize or reconstruct
the state space, and then identify the resulting one-particle spectral
subspaces.  Counting component fields in an off-shell superfield is therefore
not a proof of the number or spin content of physical particles.

\paragraph{On-shell particle multiplets.}
Let \(\mathcal V\) be a finite-dimensional linear space of component field
variables on which the supertranslation algebra acts by linear differential
operators in spacetime, and let \(S_{\rm cl}\) be a quadratic supersymmetric
action.  The physical one-particle multiplet is determined by the
positive-frequency solution space of the Euler--Lagrange equations modulo
linearized gauge transformations and null vectors of the symplectic form; it
is not determined by \(\mathcal V\) alone.  In particular, an auxiliary field
whose equation of motion is algebraic contributes no independent one-particle
oscillator in the free theory.
The free one-particle space is obtained from the space of classical solutions
by the standard linear quantization construction: one restricts to solutions
of the linear equations, quotients pure-gauge directions if present, equips
the quotient with the conserved Hermitian form coming from the quadratic
action, and completes the positive-frequency part.  None of these operations
is encoded in the representation of the supertranslation algebra on the
ambient coordinate space \(\mathcal V\) alone.

If a component \(F\in\mathcal V\) is auxiliary, the quadratic action contains
no time derivative of \(F\) and the Euler--Lagrange equation has the form
\[
  A F+B(\Phi)=0,
\]
where \(A\) is an invertible algebraic operator and \(\Phi\) denotes the other
linearized fields.  Thus \(F=-A^{-1}B(\Phi)\) on the solution space.  The
initial data of \(F\) are fixed by the initial data of \(\Phi\), and the
symplectic form has no independent \(F\)-oscillator.  Consequently the
auxiliary component may be necessary for off-shell closure, but it is absent
from the particle Hilbert-space multiplet.

\section{Graded Lie Algebra}

\begin{definition}[Four-dimensional \(\mathcal N=1\) supertranslation algebra]
\label{def:n1-supertranslation-algebra}
Let \(P_\mu\) be the translation generators in the convention
\(U(a)=\exp(\ii a^\mu P_\mu)\).  The complexified \(\mathcal N=1\)
supertranslation algebra is generated by \(P_\mu\), odd Weyl spinors
\(Q_\alpha\), \(\bar Q_{\dot\alpha}\), and the identity, with
\[
  [P_\mu,P_\nu]=[P_\mu,Q_\alpha]=[P_\mu,\bar Q_{\dot\alpha}]=0,
\]
\[
  \{Q_\alpha,Q_\beta\}=0,
  \qquad
  \{\bar Q_{\dot\alpha},\bar Q_{\dot\beta}\}=0,
  \qquad
  \{Q_\alpha,\bar Q_{\dot\beta}\}
  =
  2\sigma^\mu_{\alpha\dot\beta}P_\mu .
\]
The matrices \(\sigma^\mu_{\alpha\dot\beta}\) and the raising/lowering
conventions are those of Section~\ref{sec:spinor-conventions}.  A real form is obtained by a
choice of adjoint operation for which \(P_\mu\) is Hermitian and
\((Q_\alpha)^\dagger=\bar Q_{\dot\alpha}\) after the stated spinor
identifications.
\end{definition}

The Poincare generators \(M_{\mu\nu}\) act on \(Q\) and \(\bar Q\) by the two
Weyl spinor representations.  This action is part of the full super-Poincare
algebra and fixes the Lorentz covariance of the anticommutator above.

\section{Haag--\L opusza\'nski--Sohnius Classification}

The super-Poincare algebra is not introduced as an arbitrary formal gadget.
Its distinguished status in four-dimensional particle physics comes from a
classification theorem about possible symmetries of a massive scattering
theory.  The theorem is conditional.  It assumes the same kind of scattering
framework as the Coleman--Mandula theorem: a positive-energy Hilbert space,
an \(S\)-matrix on an asymptotic particle space, analyticity of matrix
elements in the relevant variables, a finite number of one-particle species
below each mass, and nontrivial scattering.  These assumptions exclude
massless conformal theories without an ordinary particle \(S\)-matrix,
integrable two-dimensional theories with infinitely many higher-spin charges,
and gauge-theory charged sectors whose asymptotic states require nonlocal
dressings.  They are therefore hypotheses of a classification result, not
axioms of QFT.

\begin{definition}[Finite graded scattering symmetry]
\label{def:finite-graded-scattering-symmetry}
In the setting of a massive Haag--Ruelle scattering theory, a finite graded
scattering symmetry is a finite-dimensional real Lie superalgebra
\[
  \mathfrak g=\mathfrak g_{\bar 0}\oplus \mathfrak g_{\bar 1}
\]
represented by densely defined operators on a common invariant domain in the
asymptotic Hilbert space, with the following properties.
\begin{enumerate}
  \item The even subalgebra contains the Poincare algebra generated by
  \(P_\mu\) and \(M_{\mu\nu}\).
  \item The even generators preserve fermion parity and the odd generators
  reverse it.
  \item The generators commute with the scattering operator \(S\) on the
  common domain, after acting by the corresponding graded Leibniz rule on
  tensor-product asymptotic states.
  \item The action on one-particle states has finite multiplicity at fixed
  mass and spin, and no generator creates a continuum of new particle species
  from one species.
\end{enumerate}
The word ``finite'' refers to the symmetry algebra and to particle-species
multiplicity; it does not assert that the Hilbert space is finite-dimensional.
\end{definition}

\begin{quotedtheorem}[Haag--\L opusza\'nski--Sohnius form in four dimensions]
\label{qt:hls-four-dimensional-form}
Under the usual analyticity, finite-species, positive-energy, and nontrivial
massive scattering hypotheses of the Coleman--Mandula and
Haag--\L opusza\'nski--Sohnius theorems, any finite graded scattering symmetry
of a four-dimensional theory is, after quotienting possible ideals acting
trivially on the \(S\)-matrix, a super-Poincare algebra with compact internal
even symmetries.  More explicitly, the even generators are
\[
  P_\mu,\qquad M_{\mu\nu},\qquad B_A ,
\]
where \(B_A\) generate a compact internal Lie algebra commuting with
\(P_\mu\) and \(M_{\mu\nu}\).  The odd generators are Weyl spinors
\[
  Q^I_\alpha,\qquad \bar Q_{\dot\alpha I},
  \qquad
  I=1,\ldots,\mathcal N,\quad \alpha,\dot\alpha=1,2,
\]
possibly transforming in a unitary representation of the internal algebra.
In a basis compatible with the positive inner product, the nonzero brackets
may be written
\begin{align}
  [M_{\mu\nu},Q^I_\alpha]
  &=
  (\sigma_{\mu\nu})_\alpha{}^\beta Q^I_\beta,
  &
  [M_{\mu\nu},\bar Q_{\dot\alpha I}]
  &=
  (\bar\sigma_{\mu\nu})_{\dot\alpha}{}^{\dot\beta}
  \bar Q_{\dot\beta I},
  \label{eq:hls-lorentz-action}
  \\
  [B_A,Q^I_\alpha]
  &=
  (T_A)^I{}_J Q^J_\alpha,
  &
  [B_A,\bar Q_{\dot\alpha I}]
  &=
  -\bar Q_{\dot\alpha J}(T_A)^J{}_I,
  \label{eq:hls-internal-action}
  \\
  \{Q^I_\alpha,\bar Q_{\dot\beta J}\}
  &=
  2\delta^I{}_J\sigma^\mu_{\alpha\dot\beta}P_\mu,
  \label{eq:hls-q-qbar}
  \\
  \{Q^I_\alpha,Q^J_\beta\}
  &=
  2\epsilon_{\alpha\beta} Z^{IJ},
  \qquad
  \{\bar Q_{\dot\alpha I},\bar Q_{\dot\beta J}\}
  =
  2\epsilon_{\dot\alpha\dot\beta}\bar Z_{IJ}.
  \label{eq:hls-central-charges}
\end{align}
Here \(Z^{IJ}=-Z^{JI}\) are central charges invariant under whatever internal
symmetry is kept as a simultaneous symmetry of the fixed algebra:
\begin{equation}
  (T_A)^I{}_K Z^{KJ}+(T_A)^J{}_K Z^{IK}=0 .
  \label{eq:hls-central-invariance}
\end{equation}
No odd Lorentz tensors of spin \(3/2\) or higher, and no additional even
spacetime-tensor charges, occur under these hypotheses.
\end{quotedtheorem}

\begin{remark}[Status of the theorem in this monograph]
\label{rem:hls-status}
The quoted theorem is a classification theorem about symmetries of a class of
massive scattering theories.  It is not a construction theorem for
supersymmetric QFTs.  It also does not apply without modification to
superconformal theories, theories without an ordinary particle \(S\)-matrix,
two-dimensional factorized theories with infinitely many higher-spin charges,
or gauge-theory charged sectors requiring Wilson-line dressings.  The
algebraic consequences used below, however, are finite-dimensional
representation-theoretic statements and are proved directly.
\end{remark}

\paragraph{Lorentz covariance of the mixed supercharge anticommutator.}
Assume the odd generators transform as left- and right-handed Weyl spinors
and commute with translations.  Under the HLS scattering hypotheses excluding
new independent even spacetime-vector charges, Lorentz covariance and
positivity put the mixed anticommutator in the form
\[
  \{Q^I_\alpha,\bar Q_{\dot\beta J}\}
  =
  2\delta^I{}_J\sigma^\mu_{\alpha\dot\beta}P_\mu
\]
after a nonsingular change of basis in the index \(I\).
The tensor product of the left Weyl representation \((\frac12,0)\) with the
right Weyl representation \((0,\frac12)\) is the vector representation
\((\frac12,\frac12)\).  Therefore Lorentz covariance requires
\[
  \{Q^I_\alpha,\bar Q_{\dot\beta J}\}
  =
  A^I{}_J\sigma^\mu_{\alpha\dot\beta}V_\mu
\]
for some even vector generator \(V_\mu\), up to possible sums of such vector
generators.  The Coleman--Mandula part of the HLS hypotheses rules out a new
independent even vector symmetry acting nontrivially on the \(S\)-matrix;
the only available spacetime vector generator is \(P_\mu\).  Thus
\[
  \{Q^I_\alpha,\bar Q_{\dot\beta J}\}
  =
  A^I{}_J\sigma^\mu_{\alpha\dot\beta}P_\mu .
\]

Hermitian conjugation gives \(A=A^\dagger\).  In a timelike positive-energy
state with momentum \(p\), the expectation value of
\(\{u_I^\alpha Q^I_\alpha,(u_J^\beta Q^J_\beta)^\dagger\}\) is a norm and is
nonnegative for every \(u\).  Hence \(A\) is positive semidefinite on the
subspace of nontrivially acting supercharges.  Quotienting zero-norm
supercharges and choosing an orthonormal basis in the remaining internal
index diagonalizes \(A\) to a positive diagonal matrix.  Rescaling each
nonzero \(Q^I\) fixes the diagonal entries to \(2\), giving the displayed
normalization.

\paragraph{Central-charge tensor structure.}
Under the same HLS hypotheses, the anticommutator of two left-handed
supercharges has only the Lorentz-scalar central part
\[
  \{Q^I_\alpha,Q^J_\beta\}
  =
  2\epsilon_{\alpha\beta}Z^{IJ},
  \qquad Z^{IJ}=-Z^{JI},
\]
with the conjugate formula for two right-handed supercharges.  If an internal
generator \(B_A\) is retained as an actual symmetry of the same algebra, then
\(Z\) obeys the invariance condition
\eqref{eq:hls-central-invariance}.

The product of two left Weyl representations decomposes as
\[
  \left(\frac12,0\right)\otimes
  \left(\frac12,0\right)
  =
  (0,0)\oplus (1,0).
\]
The scalar summand is obtained by contracting the spinor indices with
\(\epsilon_{\alpha\beta}\), while the \((1,0)\) summand is a self-dual
two-form spin representation.  An even self-dual tensor charge would be an
additional spacetime-tensor symmetry; the HLS hypotheses exclude such charges
in the particle \(S\)-matrix setting.  The remaining scalar term is therefore
\(2\epsilon_{\alpha\beta}Z^{IJ}\).  The factor \(2\) is part of the central
charge convention used below: it makes the rest-frame positivity block have
eigenvalues \(2(m\pm z_s)\) when \(z_s\) is a singular value of \(Z\).

The graded anticommutator is symmetric under exchange of the combined labels
\((I,\alpha)\leftrightarrow(J,\beta)\).  Since
\(\epsilon_{\beta\alpha}=-\epsilon_{\alpha\beta}\), this symmetry requires
\[
  2\epsilon_{\alpha\beta}Z^{IJ}
  =
  2\epsilon_{\beta\alpha}Z^{JI}
  =
  -2\epsilon_{\alpha\beta}Z^{JI},
\]
and hence \(Z^{IJ}=-Z^{JI}\).  Finally, apply the graded Jacobi identity to
\((B_A,Q^I_\alpha,Q^J_\beta)\).  Using
\([B_A,Q^I_\alpha]=(T_A)^I{}_KQ^K_\alpha\), the Jacobi identity gives
\[
  [B_A,Z^{IJ}]
  =
  (T_A)^I{}_K Z^{KJ}+(T_A)^J{}_K Z^{IK}.
\]
For \(Z^{IJ}\) to be a numerical central element in the algebra, the left
side vanishes; this is exactly \eqref{eq:hls-central-invariance}.  If this
condition fails for a proposed internal generator, that generator is an
automorphism of a family of algebras or is broken in the sector with the fixed
central charge, not a simultaneous internal symmetry of that fixed algebra.

\section{Positivity and Multiplet Size}

For a one-particle state of mass \(m>0\), go to its rest frame
\(P^\mu=(m,0,0,0)\).  The supercharge anticommutator becomes
\[
  \{Q_\alpha,Q_\beta^\dagger\}=2m\,\delta_{\alpha\beta}
\]
after choosing the rest-frame spinor basis.  Thus
\[
  a_\alpha=\frac{1}{\sqrt{2m}}Q_\alpha,\qquad
  a_\alpha^\dagger=\frac{1}{\sqrt{2m}}Q_\alpha^\dagger
\]
generate a two-mode fermionic oscillator algebra.

\paragraph{Massive \(\mathcal N=1\) oscillator module.}
Fix a massive one-particle orbit of mass \(m>0\), and fix an irreducible
spin-\(j\) representation \(V_j\) of the massive little group \(SU(2)\).  The
corresponding irreducible rest-frame supertranslation module is, as a
\(\mathbb Z_2\)-graded vector space,
\[
  V_j\otimes \Lambda^\bullet \mathbb C^2 .
\]
It has total dimension \(4(2j+1)\), bosonic dimension \(2(2j+1)\), and
fermionic dimension \(2(2j+1)\).  Under the diagonal little-group action the
fermionic one-oscillator sector carries
\begin{equation}
  V_j\otimes \mathbb C^2
  \cong
  V_{j+1/2}\oplus V_{j-1/2},
  \label{eq:massive-n1-spin-decomposition}
\end{equation}
with the \(V_{j-1/2}\) summand omitted when \(j=0\), while the zero- and
two-oscillator sectors are two copies of \(V_j\) with opposite fermion
number parity from the one-oscillator sector.
The normalized operators \(a_\alpha,a_\alpha^\dagger\) satisfy
\[
  \{a_\alpha,a_\beta^\dagger\}=\delta_{\alpha\beta},
  \qquad
  \{a_\alpha,a_\beta\}=\{a_\alpha^\dagger,a_\beta^\dagger\}=0 .
\]
The irreducible representation of this Clifford algebra with a cyclic vector
annihilated by the \(a_\alpha\) is the exterior algebra
\(\Lambda^\bullet\mathbb C^2\).  Tensoring with the chosen little-group
representation gives the stated vector space and dimension.  The two
creation operators transform as a spin-\(1/2\) doublet, so the
one-oscillator sector is \(V_j\otimes V_{1/2}\), which decomposes by the
Clebsch--Gordan rule into \eqref{eq:massive-n1-spin-decomposition}.  The
zero-oscillator sector \(\Lambda^0\mathbb C^2\) and the two-oscillator sector
\(\Lambda^2\mathbb C^2\) are both little-group scalars, giving two copies of
\(V_j\).  Fermion parity is exterior degree modulo \(2\), hence the even and
odd dimensions are both \(2(2j+1)\).

\begin{remark}[Equal bosonic and fermionic dimensions]
\label{rem:susy-equal-boson-fermion}
Let \(F=(-1)^{N_F}\) be the fermion-parity operator of the massive rest-frame
\(\mathcal N=1\) oscillator algebra.
For a two-mode fermionic Fock space,
\[
  \operatorname{Tr}F=(1-1)^2=0 .
\]
Tensoring with a little-group representation multiplies the trace by its
dimension, so the trace remains zero.  Since \(F\) has eigenvalue \(+1\) on
bosonic states and \(-1\) on fermionic states, \(\operatorname{Tr}F=0\) is
exactly equality of the two dimensions.
\end{remark}

\section{Massless Representations}

For a massless momentum \(p^\mu=(E,0,0,E)\), the matrix
\(\sigma^\mu p_\mu\) has rank one.  Only one complex linear combination of
the supercharges acts with nonzero norm on a positive-energy representation;
the null combination is quotiented.  The remaining creation operator changes
helicity by \(1/2\).  Hence a massless irreducible representation is built
from a Clifford algebra with one fermionic oscillator, and its helicities
come in adjacent pairs \((h,h+1/2)\) before CPT completion.

\paragraph{Massless \(\mathcal N=1\) helicity pair.}
Let \(E>0\) and \(p^\mu=(E,0,0,E)\).  In a spinor basis where
\(\sigma^\mu p_\mu\) is diagonal, positivity sets one complex supercharge
combination to zero on the Hilbert space and leaves one normalized
fermionic oscillator.  An irreducible massless \(\mathcal N=1\)
representation generated from a helicity-\(h\) state therefore has two
helicity states,
\[
  h,\qquad h+\frac12,
\]
before imposing CPT.
For the chosen momentum, \(\sigma^\mu p_\mu\) is a positive semidefinite
\(2\times2\) Hermitian matrix of rank one.  A unitary change of spinor basis
therefore puts the supercharge norm matrix into the form
\[
  \{Q_a,Q_b^\dagger\}
  =
  \begin{pmatrix}
    cE&0\\
    0&0
  \end{pmatrix}_{ab},
  \qquad c>0 .
\]
If \(\psi\) is any vector in the common invariant domain, then
\[
  0
  =
  \langle \psi,\{Q_2,Q_2^\dagger\}\psi\rangle
  =
  \|Q_2\psi\|^2+\|Q_2^\dagger\psi\|^2
\]
because the second diagonal entry is zero.  Positivity of the inner product
therefore forces \(Q_2\psi=Q_2^\dagger\psi=0\) for all vectors in the
domain, equivalently the null supercharge combination acts trivially after
quotienting zero-norm vectors.  The remaining normalized operator
\[
  b=(cE)^{-1/2}Q_1,\qquad b^\dagger=(cE)^{-1/2}Q_1^\dagger
\]
satisfies \(\{b,b^\dagger\}=1\).  Because a Weyl supercharge has helicity
\(\pm1/2\), the nontrivial creation operator shifts helicity by one half in
the chosen convention.  Thus the irreducible Clifford module has precisely
the two displayed helicities.

\section{Central Charges and BPS Bounds}

For extended supersymmetry, odd generators carry an index \(I=1,\ldots,N\).
The algebra may contain antisymmetric central charges,
\[
  \{Q^I_\alpha,Q^J_\beta\}
  =
  2\epsilon_{\alpha\beta}Z^{IJ},
  \qquad Z^{IJ}=-Z^{JI}.
\]
In a rest frame, positivity of the matrix of anticommutators gives lower
bounds on the mass in terms of the singular values of \(Z\).  A state
saturating such a bound is annihilated by a nonzero linear combination of
supercharges; its multiplet is shorter because the corresponding fermionic
oscillator is absent in the quotient representation.

\begin{proposition}[Rest-frame BPS bound from singular values]
\label{prop:bps-bound-singular-values}
In a massive rest frame for extended supersymmetry, suppose the central-charge
matrix has singular values \(z_s\ge0\) after skew-diagonalization of the
antisymmetric matrix \(Z^{IJ}\).  Positivity of the Hilbert-space
representation implies
\[
  m\ge z_s
  \qquad\text{for every singular value }z_s .
\]
If \(m=z_s\) for some \(s\), then a nonzero linear combination of
supercharges has zero norm and annihilates the shortened representation.
\end{proposition}

\begin{proof}
The linear-algebra input is the skew Takagi factorization for a complex
antisymmetric matrix: there is a unitary matrix \(U\) such that
\[
  (U Z U^T)^{2s-1,2s}=z_s,\qquad
  (U Z U^T)^{2s,2s-1}=-z_s,\qquad z_s\ge0,
\]
with all other entries inside the displayed \(2\times2\) block zero, and with
additional zero blocks if the number of supercharges is odd.  Replacing
\(Q^I_\alpha\) by \(U^I{}_JQ^J_\alpha\) preserves the mixed anticommutator
\[
  \{Q^I_\alpha,(Q^J_\beta)^\dagger\}=2m\,\delta_{\alpha\beta}\delta^{IJ}
\]
in the rest frame and puts the central term into these blocks.

Fix one nonzero block and write it as \(I=1,2\), \(Z^{12}=z\).  Since
\((Q^I_\alpha)^\dagger=\bar Q_{\dot\alpha I}\), the two-dimensional subspace
spanned by the operators \(Q^1_1\) and \(\bar Q_{\dot2\,2}\) has Hermitian
anticommutator matrix
\[
  \begin{pmatrix}
  \{Q^1_1,\bar Q_{\dot1\,1}\}
  &
  \{Q^1_1,Q^2_2\}
  \\
  \{\bar Q_{\dot2\,2},\bar Q_{\dot1\,1}\}
  &
  \{\bar Q_{\dot2\,2},Q^2_2\}
  \end{pmatrix}
  =
  2
  \begin{pmatrix}
  m & z\\
  z & m
  \end{pmatrix},
\]
after the phase of the block has been absorbed into the unitary Takagi basis.
The companion pair \(Q^1_2,-\bar Q_{\dot1\,2}\) gives the same matrix.  Thus
the normalized combinations
\[
  R_\pm=\frac{1}{\sqrt2}\left(Q^1_1\pm \bar Q_{\dot2\,2}\right)
\]
have anticommutators
\[
  \{R_\pm,R_\pm^\dagger\}=2(m\pm z).
\]
The same computation applies independently to every skew block and to each
multiplicity space for a repeated singular value.

In a Hilbert-space representation, \(\langle\psi,\{R,R^\dagger\}\psi\rangle
=\|R^\dagger\psi\|^2+\|R\psi\|^2\) is nonnegative for every vector in the
common invariant domain of the supercharges.  Therefore every eigenvalue of
each block must be nonnegative, and in particular \(2(m-z_s)\ge0\).  This gives
\(m\ge z_s\) for every singular value.  If \(m=z_s\), the operator \(R_-\) for
that block has zero norm on every vector; after quotienting zero-norm vectors
and closing the operator, \(R_-\) acts trivially.  The associated fermionic
oscillator is absent, which is the representation-theoretic shortening
mechanism.  This proves only the algebraic BPS inequality and shortening
criterion; whether a QFT contains states saturating it is separate dynamical
information.
\end{proof}

\begin{openproblem}[Nonperturbative realization of supersymmetric QFT data]
\label{op:susy-nonperturbative-realization}
For a proposed supersymmetric QFT, one must construct the Hilbert space or
Euclidean/functorial object, the supercharges as operators or Ward
identities, the stress tensor and supercurrent, and the required positivity
structure.  Perturbative superfield expressions and protected-index
calculations are powerful coordinates on such data, but a nonperturbative
construction requires separate control of the regulator, local operator
domains, anomalies, and possible supersymmetry-breaking effects.
\end{openproblem}
