\chapter{Six-Dimensional Superconformal Theories}
\label{ch:susy-six-dimensional-scft}

\ConventionsMixed

Six-dimensional superconformal theories are QFTs whose central examples do
not admit ordinary Lagrangian descriptions at the fixed point.  They
therefore force a separation between the definition of a theory and the use
of effective field variables on a branch of vacua.  This chapter records the
objects that can be defined without assuming a microscopic Lagrangian:
superconformal representation data, tensor-branch effective actions, anomaly
polynomials, BPS strings, compactification tests, and extended operators.
The emphasis is the \((2,0)\) family and the adjacent \((1,0)\) tensor-branch
formalism, because these are the supersymmetric field theories that most
sharply test the distinction between protected data and construction of a
local QFT.

\section{Six-Dimensional Supersymmetry Data}

Let the Lorentz group be \(\operatorname{Spin}(1,5)\).  A chiral spinor has
four complex components before imposing a symplectic-Majorana-Weyl reality
condition.  Six-dimensional supersymmetry is labeled by
\((\mathcal N_+,\mathcal N_-)\), the numbers of positive and negative
chirality symplectic-Majorana-Weyl supercharges.  The cases most central for
interacting local fixed points are \((1,0)\) and \((2,0)\).

For \((1,0)\), the R-symmetry is \(SU(2)_R\).  For \((2,0)\), the
R-symmetry is \(USp(4)_R\simeq \operatorname{Spin}(5)_R\).  Superconformal
symmetry enlarges the bosonic group to
\[
  SO(2,6)\times G_R ,
\]
with \(G_R=SU(2)_R\) or \(USp(4)_R\).  The full superconformal algebras are
usually denoted \(OSp(8^\ast|2)\) for \((1,0)\) and \(OSp(8^\ast|4)\) for
\((2,0)\).  Local operators are organized as positive-energy
representations of these algebras on the cylinder.  This statement concerns
Hilbert-space representation theory; it does not supply off-shell fields.

\section{Why Six-Dimensional Gauge Fields Are Branch Variables}

A six-dimensional vector multiplet can appear in an effective gauge-theory
description on a tensor branch, but it is not a conformal gauge field at an
interacting fixed point.

\begin{remark}[Dimension of the Yang--Mills coupling]
\label{rem:six-d-yang-mills-coupling-dimension}
In a six-dimensional Yang--Mills presentation with canonical gauge field
normalization, the gauge coupling has mass dimension \(-1\), equivalently
\([g_6^2]=-2\).  Hence the local Yang--Mills interaction is not a marginal
microscopic coupling of a six-dimensional conformal field theory.

In \(d\) spacetime dimensions a canonically normalized vector field has mass
dimension
\[
  [A_\mu]=\frac{d-2}{2}.
\]
The covariant derivative is \(\partial_\mu+\ii g_d A_\mu\), so
\[
  [g_d]+[A_\mu]=1 .
\]
Therefore
\[
  [g_d]=1-\frac{d-2}{2}=\frac{4-d}{2},
  \qquad
  [g_d^2]=4-d .
\]
Putting \(d=6\) gives \([g_6]=-1\) and \([g_6^2]=-2\).  A coupling with
negative mass dimension is not a marginal conformal coordinate.  Thus any
six-dimensional vector-multiplet gauge theory used below is an effective
description on a branch, not a Lagrangian definition of the fixed point.
\end{remark}

\section{Multiplets And Self-Dual Fields}

A \((1,0)\) tensor multiplet contains a two-form gauge potential \(B\) with
self-dual field strength, a real scalar \(\phi\), and fermions.  A
hypermultiplet contains scalars in quaternionic representations and fermions.
A \((2,0)\) tensor multiplet contains
\[
  B,\qquad \phi^A\;(A=1,\ldots,5),\qquad \text{fermions},
\]
where the five real scalars transform as the vector representation of
\(\operatorname{Spin}(5)_R\).  The field strength
\[
  H=dB+\cdots
\]
obeys a self-duality constraint
\[
  H=*H .
\]
For a nonabelian interacting \((2,0)\) theory there is no known ordinary
local action written in terms of nonabelian two-form gauge potentials.  Even
for abelian self-dual fields, a covariant action requires auxiliary
structure or a formulation in which self-duality is imposed after variation.
Consequently this chapter states self-dual fields through their equations,
charge lattice, anomaly inflow, and compactification tests.

\begin{definition}[Abelian self-dual tensor system]
\label{def:six-d-abelian-self-dual-tensor-system}
An abelian six-dimensional tensor sector on a spacetime \(M_6\) consists,
locally, of two-form gauge potentials \(B^I\) with gauge transformations
\[
  B^I\longmapsto B^I+d\Lambda^I ,
\]
three-form field strengths \(H^I\), an integral string-charge lattice
\(\Lambda_{\rm str}\), and a nondegenerate symmetric Dirac pairing
\[
  \Omega:\Lambda_{\rm str}\times\Lambda_{\rm str}\longrightarrow\mathbb Z .
\]
In a topologically trivial local chart without background Chern--Simons
corrections, \(H^I=dB^I\).  Globally, \(B^I\) is better regarded as a
differential-cohomology field, and the periods of \(H^I/2\pi\) pair
integrally with string charges.  A self-dual tensor field is not specified by
an unconstrained Gaussian integral over \(B^I\).  Its local equations include
a chirality constraint of the form
\[
  H^I = *H^I
\]
after choosing the self-dual subspace determined by the metric and the
pairing convention.  If both self-dual and anti-self-dual tensors are present,
the chirality matrix is correspondingly part of the system.
\end{definition}

\paragraph{On-shell degrees of freedom of a chiral two-form.}
For a free massless two-form gauge potential in six-dimensional Minkowski
space, the gauge-invariant one-particle polarizations before imposing
self-duality form \(\Lambda^2\mathbb R^4\), where \(\mathbb R^4\) is the
transverse little-group vector representation.  Thus an ordinary two-form has
\(\binom42=6\) physical bosonic polarizations.  The Hodge star on
\(\Lambda^2\mathbb R^4\) squares to \(1\), so a chiral two-form has
\[
  \frac12\binom42=3
\]
physical bosonic polarizations.  A free \((2,0)\) tensor multiplet therefore
has \(3+5=8\) bosonic on-shell polarizations.

For a massless particle choose momentum \(k\) and quotient by gauge
transformations \(B\mapsto B+k\wedge\lambda\).  The physical polarizations
are transverse two-forms modulo longitudinal ones; equivalently they are
two-forms on the \(D-2=4\) dimensional transverse representation of the
massless little group \(SO(4)\).  This gives \(\binom42=6\) polarizations for
an unconstrained two-form.  The six-dimensional self-duality condition on
the field strength reduces, on the one-particle transverse space, to the
four-dimensional decomposition
\[
  \Lambda^2\mathbb R^4
  =
  \Lambda^2_+\mathbb R^4\oplus
  \Lambda^2_-\mathbb R^4 ,
  \qquad
  \dim\Lambda^2_\pm\mathbb R^4=3 .
\]
Choosing one chirality leaves three polarizations.  The five scalars of the
\((2,0)\) tensor multiplet add five further bosonic polarizations, for a
total of eight.

\section{Tensor-Branch Datum}

On a tensor branch, expectation values of tensor-multiplet scalars define
coordinates
\[
  \phi^I\in\mathcal C ,
\]
where \(\mathcal C\) is a cone modulo any discrete identifications.  The
low-energy action contains
\begin{equation}
  \frac12 G_{IJ}\,d\phi^I\wedge *d\phi^J
  +
  \frac12 G_{IJ}\,H^I\wedge *H^J
  +
  B^I\wedge X_I^{(4)}
  +\cdots ,
  \label{eq:six-d-tensor-branch-action}
\end{equation}
with self-duality imposed as a constraint on the equations of motion or by a
chosen self-dual-field formalism.  The four-form \(X_I^{(4)}\) encodes gauge,
flavor, and gravitational couplings of the branch EFT.

\begin{definition}[Tensor-branch effective theory]
\label{def:six-d-tensor-branch-effective-theory}
A tensor-branch effective theory consists of
\[
\begin{gathered}
  \Lambda_{\rm str},\quad \Omega_{IJ},\quad
  \phi\in\mathcal C\subset \Lambda_{\rm str}^{\vee}\otimes\mathbb R^n,
  \quad
  B^I,\ H^I,\\
  X_I^{(4)}\in\Omega^4(M_6),
  \qquad
  \text{a self-dual tensor quantization prescription.}
\end{gathered}
\]
The four-forms \(X_I^{(4)}\) are built from background gauge, flavor,
R-symmetry, and tangent-bundle curvatures.  Here \(n=1\) for a \((1,0)\)
tensor scalar and \(n=5\) for the
\((2,0)\) \(\operatorname{Spin}(5)_R\)-vector scalar.  The cone
\(\mathcal C\) is divided into chambers by loci where charged strings become
tensionless.  The low-energy formulae on one chamber are not a construction
of the fixed point at the chamber wall.
\end{definition}

For a \((2,0)\) theory of simply laced type \(\mathfrak g\), with Cartan
algebra \(\mathfrak t_{\mathfrak g}\), rank \(r_{\mathfrak g}\), and Weyl
group \(W_{\mathfrak g}\), the tensor-branch scalar space has the form
\[
  \mathcal C_{\mathfrak g}^{(2,0)}
  =
  \bigl(\mathbb R^5\otimes\mathfrak t_{\mathfrak g}\bigr)/W_{\mathfrak g}
\]
after removing any decoupled center-of-mass tensor multiplet.  The five
\(\mathbb R^5\) directions are the \(\operatorname{Spin}(5)_R\) vector
directions.  The Weyl quotient records the identification of Cartan
coordinates that survive from the interacting nonabelian theory.

\begin{example}[Type \(A_{N-1}\) tensor branch]
\label{ex:six-d-a-type-tensor-branch}
For type \(A_{N-1}\), the rank is \(N-1\), and the tensor branch may be
written as
\[
  \left\{
    (x_1,\ldots,x_N)\in(\mathbb R^5)^N:
    \sum_{i=1}^N x_i=0
  \right\}/S_N .
\]
The symmetric group permutes the \(N\) eigenvalue-like vectors.  This formula
is a tensor-branch coordinate chart; it is not a definition of the
interacting fixed point at the origin \(x_i=x_j\).
\end{example}

\section{Anomaly Polynomial}

For a six-dimensional theory coupled to background metric and R-symmetry
fields, the anomaly is encoded by an eight-form
\[
  I_8
  \in
  H^8(BG_{\mathrm{background}},\mathbb Q)
\]
whose descent gives anomalous variations of the generating functional.  For a
\((1,0)\) theory one may write
\[
  I_8
  =
  \alpha\,c_2(R)^2
  +
  \beta\,c_2(R)p_1(T)
  +
  \gamma\,p_1(T)^2
  +
  \delta\,p_2(T)
  +
  I_8^{\mathrm{flavor}},
\]
where \(c_2(R)\) is the second Chern class of the \(SU(2)_R\) bundle and
\(p_i(T)\) are Pontryagin classes of the tangent bundle.  The coefficients
are part of the QFT structure.

On a tensor branch, anomaly matching takes the form
\begin{equation}
  I_8^{\mathrm{UV}}-I_8^{\mathrm{branch}}
  =
  \frac12\Omega^{IJ}X_I^{(4)}X_J^{(4)} ,
  \label{eq:six-d-gs-anomaly-matching}
\end{equation}
where \(\Omega^{IJ}\) is the inverse tensor pairing.  The right-hand side is
produced by Green--Schwarz couplings \(B^I\wedge X_I^{(4)}\).  This equation
is a computable constraint on branch EFTs and on proposed fixed points.

\paragraph{Quadratic Green--Schwarz descent calculation.}
Fix the descent convention in which an anomaly polynomial \(P_8\) contributes
\[
  \delta_\alpha \Gamma = 2\pi\ii\int_{M_6}P_6^{(1)}(\alpha)
\]
when \(P_8=dP_7^{(0)}\) and
\(\delta_\alpha P_7^{(0)}=dP_6^{(1)}(\alpha)\).  Suppose
\[
  X_I^{(4)}=dX_I^{(3)},\qquad
  \delta_\alpha X_I^{(3)}=dX_I^{(2)}(\alpha),
\]
and \(\Omega^{IJ}\) is symmetric.  Then the quadratic polynomial
\begin{equation}
  P_8^{\rm GS}
  =
  \frac12\Omega^{IJ}X_I^{(4)}X_J^{(4)}
  \label{eq:six-d-gs-quadratic-polynomial}
\end{equation}
has descent representative
\begin{equation}
  P_6^{(1),{\rm GS}}(\alpha)
  =
  \Omega^{IJ}X_I^{(2)}(\alpha)X_J^{(4)} .
  \label{eq:six-d-gs-descent-six-form}
\end{equation}
With the sign of the local Green--Schwarz functional chosen so that its
variation is \(2\pi\ii\int P_6^{(1),{\rm GS}}\), the tensor-branch matching
condition is precisely
\[
  I_8^{\rm UV}=I_8^{\rm branch}+P_8^{\rm GS}.
\]
The substantive six-dimensional input is the anomaly polynomial of the fixed
point and of the branch theory; the following calculation fixes the local
descent representative and sign convention.  Because \(X_I^{(4)}\) has even
degree, the product
\(X_I^{(4)}X_J^{(4)}\) is symmetric under \(I\leftrightarrow J\).  A descent
seven-form for \eqref{eq:six-d-gs-quadratic-polynomial} is
\[
  P_7^{(0),{\rm GS}}
  =
  \frac12\Omega^{IJ}
  \bigl(
    X_I^{(3)}X_J^{(4)}
    +
    X_J^{(3)}X_I^{(4)}
  \bigr),
\]
which equals \(\Omega^{IJ}X_I^{(3)}X_J^{(4)}\) after using the symmetry of
\(\Omega^{IJ}\).  Varying gives
\[
  \delta_\alpha P_7^{(0),{\rm GS}}
  =
  \Omega^{IJ}dX_I^{(2)}(\alpha)X_J^{(4)}
  =
  d\!\left(\Omega^{IJ}X_I^{(2)}(\alpha)X_J^{(4)}\right),
\]
since \(dX_J^{(4)}=0\).  This proves
\eqref{eq:six-d-gs-descent-six-form}.  The final matching equation is the
statement that the anomaly of the branch massless fields plus the anomaly
generated by the Green--Schwarz term equals the anomaly of the fixed point.

\begin{quotedtheorem}[\((2,0)\) anomaly polynomial]
\label{qt:six-d-two-zero-anomaly-polynomial}
Let \(\mathfrak g\) be simply laced, with rank \(r_{\mathfrak g}\), dimension
\(d_{\mathfrak g}\), and dual Coxeter number \(h^\vee_{\mathfrak g}\).  The
\((2,0)\) theory of type \(\mathfrak g\) has anomaly polynomial
\begin{equation}
  I_8[\mathfrak g]
  =
  r_{\mathfrak g} I_8(1)
  +
  \frac{d_{\mathfrak g}h^\vee_{\mathfrak g}}{24}\,
  p_2(N_R),
  \label{eq:six-d-two-zero-anomaly}
\end{equation}
where \(N_R\) is the rank-five \(SO(5)_R\) background bundle and
\begin{equation}
  I_8(1)
  =
  \frac1{48}
  \left[
    p_2(N_R)-p_2(T)
    +\frac14\bigl(p_1(T)-p_1(N_R)\bigr)^2
  \right]
  \label{eq:six-d-free-tensor-anomaly}
\end{equation}
is the anomaly polynomial of one free \((2,0)\) tensor multiplet.
\end{quotedtheorem}

The theorem is quoted because the chapter does not reproduce a construction of
the interacting six-dimensional local QFT together with a first-principles
derivation of its anomaly functional.  In this monograph it is used only as
protected QFT data: it fixes anomaly coefficients that any proposed
construction of the \((2,0)\) theory must reproduce.  External geometric or
string-theoretic derivations may motivate the formula, but they are not
premises in the argument of this chapter.

\paragraph{What the quoted theorem asserts on the tensor branch.}
The generic tensor branch of the type \(\mathfrak g\) theory contains
\(r_{\mathfrak g}\) free abelian \((2,0)\) tensor multiplets.  Therefore
\begin{equation}
  I_8[\mathfrak g]-r_{\mathfrak g}I_8(1)
  =
  \frac{d_{\mathfrak g}h^\vee_{\mathfrak g}}{24}\,p_2(N_R)
  \label{eq:six-d-two-zero-interacting-excess}
\end{equation}
is the part of the anomaly which is invisible in the free branch spectrum.
The background anomaly is an invertible field theory attached to the QFT,
not to a particular coordinate chart on its vacuum branch.  Thus the
low-energy tensor-branch description must reproduce the same anomaly
functional as the fixed point after its massless fields and its local
Green--Schwarz terms are both included.  This is the precise sense in which
the formula is tested by branch data: the branch fields give
\(r_{\mathfrak g}I_8(1)\), while the remaining coefficient is a quadratic
functional of the four-form couplings \(X_I^{(4)}\) in
\eqref{eq:six-d-gs-anomaly-matching}.  It is not an additional propagating
degree-of-freedom count.

The following finite root-system facts are the arithmetic skeleton of this
matching.  Normalize the simply laced root system
\(\Delta_{\mathfrak g}\subset\mathfrak t_{\mathfrak g}^\vee\) so every root
has squared length \(2\), and identify
\(\mathfrak t_{\mathfrak g}\simeq\mathfrak t_{\mathfrak g}^\vee\) by this
inner product.  Then
\begin{equation}
  \sum_{\alpha\in\Delta_{\mathfrak g}^+}
  \alpha\otimes\alpha
  =
  h^\vee_{\mathfrak g}\,\id_{\mathfrak t_{\mathfrak g}},
  \label{eq:six-d-positive-root-second-moment}
\end{equation}
or, equivalently,
\[
  \sum_{\alpha\in\Delta_{\mathfrak g}^+}
  \langle\alpha,v\rangle^2
  =
  h^\vee_{\mathfrak g}\,\langle v,v\rangle
  \qquad
  (v\in\mathfrak t_{\mathfrak g}).
\]
Indeed the full sum over all roots defines a Weyl-invariant symmetric
endomorphism of the irreducible reflection representation of \(W_{\mathfrak g}\),
hence is a scalar multiple of the identity.  Taking the trace fixes the
scalar: the trace of the full sum is
\(\sum_{\alpha\in\Delta_{\mathfrak g}}\langle\alpha,\alpha\rangle
=2|\Delta_{\mathfrak g}|\), while
\(|\Delta_{\mathfrak g}|=d_{\mathfrak g}-r_{\mathfrak g}
=r_{\mathfrak g}h^\vee_{\mathfrak g}\).  Dividing by
\(r_{\mathfrak g}\) gives \(2h^\vee_{\mathfrak g}\) for the full root sum
and \(h^\vee_{\mathfrak g}\) for the positive-root sum.  The identity
\begin{equation}
  d_{\mathfrak g}=r_{\mathfrak g}(h^\vee_{\mathfrak g}+1)
  \label{eq:six-d-dimension-rank-coxeter-relation}
\end{equation}
is the same root count with the Cartan directions restored.

These identities do not prove
\eqref{eq:six-d-two-zero-anomaly}; they explain why, once the normalization
of the self-dual tensor Green--Schwarz coupling is fixed on a long-root
string, the Weyl-invariant quadratic contribution on a chamber has no
remaining independent symmetric tensor structure.  The nontrivial QFT datum
in the quoted theorem is the overall coefficient with which the interacting
fixed point realizes this rigid invariant.

\begin{remark}[ADE arithmetic in the \((2,0)\) anomaly coefficient]
\label{rem:six-d-ade-anomaly-arithmetic}
For simply laced \(\mathfrak g\), the coefficient multiplying
\(p_2(N_R)\) in \eqref{eq:six-d-two-zero-anomaly} is
\[
  \frac{d_{\mathfrak g}h^\vee_{\mathfrak g}}{24}.
\]
For the ADE series this gives
\[
\begin{array}{c|c|c|c|c}
  \mathfrak g & r_{\mathfrak g} & d_{\mathfrak g}
  & h^\vee_{\mathfrak g}
  & d_{\mathfrak g}h^\vee_{\mathfrak g}/24\\ \hline
  A_{N-1} & N-1 & N^2-1 & N & N(N^2-1)/24\\
  D_N & N & N(2N-1) & 2N-2 & N(N-1)(2N-1)/12\\
  E_6 & 6 & 78 & 12 & 39\\
  E_7 & 7 & 133 & 18 & 399/4\\
  E_8 & 8 & 248 & 30 & 310
\end{array}
\]
The \(A_{N-1}\) coefficient grows as \(N^3/24\) at large \(N\), while the
number of free tensor multiplets on the generic branch is only \(N-1\).

The entries use the standard simply laced Lie-algebra dimensions and dual
Coxeter numbers:
\[
  d_{A_{N-1}}=N^2-1,\quad h^\vee_{A_{N-1}}=N,
\]
\[
  d_{D_N}=N(2N-1),\quad h^\vee_{D_N}=2N-2,
\]
and
\[
  (d_{E_6},h^\vee_{E_6})=(78,12),\quad
  (d_{E_7},h^\vee_{E_7})=(133,18),\quad
  (d_{E_8},h^\vee_{E_8})=(248,30).
\]
Substitution into \(d_{\mathfrak g}h^\vee_{\mathfrak g}/24\) gives the table.
For \(A_{N-1}\),
\[
  \frac{N(N^2-1)}{24}
  =
  \frac{N^3}{24}-\frac{N}{24},
\]
which displays the cubic scaling.
\end{remark}

\section{BPS Strings}

Tensor multiplets couple electrically to strings.  A string with charge
\[
  q\in\Lambda_{\mathrm{string}}
\]
has tension
\[
  T_q=q_I\phi^I
\]
on a \((1,0)\) tensor branch, in a chamber where this pairing is positive.
For a \((2,0)\) theory of type \(\mathfrak g\), the charge lattice is
identified with the root lattice in the simply connected convention, and a
root \(\alpha\) gives a string with tension
\[
  T_\alpha=\left|\alpha(\phi)\right|,
  \qquad
  \phi\in\mathbb R^5\otimes\mathfrak t_{\mathfrak g}.
\]
At the superconformal point some tensions vanish, and the corresponding
degrees of freedom are strings, not particles in six dimensions.

\begin{proposition}[BPS string tension from the tensor-branch central charge]
\label{prop:six-d-bps-string-tension-central-charge}
Normalize the tensor-branch scalar coordinates so that the string central
charge vector of a charge \(q\in\Lambda_{\rm str}\) is
\[
  Z_q^A=q_I\phi^{IA},\qquad A=1,\ldots,n,
\]
where \(n=1\) in \((1,0)\) notation and \(n=5\) in \((2,0)\) notation.  The
positive-energy representation of the six-dimensional supersymmetry algebra
then gives the BPS inequality
\[
  T_q\geq |Z_q|,
\]
and a BPS string saturates it:
\[
  T_q=|q_I\phi^I|.
\]
For a \((2,0)\) theory of type \(\mathfrak g\), taking
\(\Lambda_{\rm str}\) to be the root lattice gives
\[
  T_\alpha=|\alpha(\phi)|
\]
for the string of root charge \(\alpha\).
\end{proposition}

\begin{proof}
Work per unit length for a straight static string oriented, say, along
\(x^5\).  Translation invariance along the string replaces the energy by a
tension \(T_q\), and the tensor-multiplet central charge appears in the
supercharge anticommutator as an R-symmetry vector \(Z_q^A\).  Choose the
positive normalization of the supercharges so that the coefficient of the
tension term in the anticommutator is \(T_q\).  After choosing a unit vector
\(\widehat Z^A=Z_q^A/|Z_q|\), the relevant Hermitian anticommutator on the
subspace of spinor/R-symmetry parameters is
\[
  \{Q(\varepsilon),Q(\varepsilon)^\dagger\}
  =
  \varepsilon^\dagger
  \bigl(T_q\,{\bf 1}-|Z_q|\,\Gamma_{\rm str}\Gamma_{\widehat Z}\bigr)
  \varepsilon .
\]
Here \(\Gamma_{\rm str}\) is the chirality matrix determined by the oriented
two-plane swept out by the static string, and \(\Gamma_{\widehat Z}\) is the
unit R-symmetry Clifford multiplication in the direction \(\widehat Z\).  The
two matrices square to \(1\) on the relevant spinor space and their product
has eigenvalues \(\pm1\).  Positivity of the Hilbert-space norm of
\(Q(\varepsilon)^\dagger|\text{string}\rangle\) for every \(\varepsilon\)
therefore requires
\[
  T_q-|Z_q|\ge0 .
\]
If equality holds, the eigenspace with
\(\Gamma_{\rm str}\Gamma_{\widehat Z}=+1\) is annihilated by a nonzero
linear combination of supercharges; conversely, preservation of a nonzero
supercharge forces the smallest eigenvalue of the displayed matrix to
vanish.  This is the BPS saturation condition.  With the scalar-coordinate
normalization in the statement, \(Z_q^A=q_I\phi^{IA}\), hence
\(T_q=|Z_q|=|q_I\phi^I|\) for a saturated string.  In the \((2,0)\) ADE
theory the charged strings on a generic tensor branch are labeled by roots,
so \(q=\alpha\) gives \(T_\alpha=|\alpha(\phi)|\).
\end{proof}

\paragraph{Local anomaly inflow to a charged string.}
Let \(\iota:\Sigma\hookrightarrow M_6\) be the oriented two-dimensional
support of a string of charge \(q\).  Write \(N_\Sigma\) for its oriented
rank-four normal bundle, and put
\[
  q^2=q_I\Omega^{IJ}q_J .
\]
The charge convention is the one used in the tensor-branch coupling:
the surface insertion contains the differential-cohomology holonomy
\[
  \exp\!\left(2\pi\ii\,q_I\int_\Sigma B^I\right).
\]
In a local background chart with \(X_I^{(4)}=dX_I^{(3)}\) and
\(\delta X_I^{(3)}=dX_I^{(2)}\), Green--Schwarz gauge invariance is expressed
by the local transformation
\[
  \delta B^I=-X_I^{(2)} .
\]
Consequently the charged insertion varies by
\[
  \delta\log W_q(\Sigma)
  =
  -2\pi\ii\int_\Sigma q_I\,\iota^\ast X_I^{(2)} .
\]
A two-dimensional quantum theory supported on \(\Sigma\) cancels this
variation precisely when its anomaly polynomial contains the four-form
\begin{equation}
  I_{4,\rm lin}^{\Sigma}(q)
  =
  q_I\,\iota^\ast X_I^{(4)} .
  \label{eq:six-d-string-linear-inflow}
\end{equation}
This is the linear inflow term.  It is not optional: it is the descent of the
same \(B^I\wedge X_I^{(4)}\) coupling that appears in the six-dimensional
Green--Schwarz matching equation.

There is also a local normal-bundle inflow from the self-pairing of the
tensor field sourced by the string.  It is best stated using the global
angular-form regularization of a tubular neighborhood of \(\Sigma\).  Let
\(S(N_\Sigma)\to\Sigma\) be the unit-sphere bundle of the normal bundle.  In
the standard normalization the Chern--Simons form
\(e_3^{(0)}(N_\Sigma)\) on this sphere bundle satisfies
\[
  de_3^{(0)}(N_\Sigma)=\pi^\ast e(N_\Sigma),
  \qquad
  \delta e_3^{(0)}(N_\Sigma)=d e_2^{(1)}(N_\Sigma),
\]
where \(e(N_\Sigma)\) is the Euler class of the rank-four normal bundle.
Replacing the delta-current Poincare dual to \(\Sigma\) by the Thom form
obtained from \(e_3^{(0)}\) gives a smooth representative of the sourced
Bianchi identity near the support.  Substituting this representative into the
quadratic tensor pairing produces, under a normal-bundle gauge
transformation, the descent variation
\[
  2\pi\ii\int_\Sigma
  \frac12\,q^2\,e_2^{(1)}(N_\Sigma).
\]
Equivalently the normal-bundle part of the inflow anomaly polynomial is
\begin{equation}
  I_{4,\rm norm}^{\Sigma}(q)
  =
  \frac12\,q_I\Omega^{IJ}q_J\,e(N_\Sigma).
  \label{eq:six-d-string-normal-inflow}
\end{equation}
The coefficient \(1/2\) is the same symmetric-pairing coefficient as in the
quadratic Green--Schwarz descent: the self-interaction of one sourced tensor
field is counted once, while distinct charge components are paired
symmetrically by \(\Omega^{IJ}\).

Thus the bulk inflow polynomial for a string of charge \(q\), in these
normalizations, is
\begin{equation}
  I_{4,\rm inflow}^{\Sigma}(q)
  =
  q_I\,\iota^\ast X_I^{(4)}
  +
  \frac12\,q_I\Omega^{IJ}q_J\,e(N_\Sigma).
  \label{eq:six-d-bps-string-inflow-polynomial}
\end{equation}
The actual anomaly polynomial of the two-dimensional degrees of freedom
supported on the string must cancel this inflow after adding its intrinsic
fermion, boson, and zero-mode contributions.  Formula
\eqref{eq:six-d-bps-string-inflow-polynomial} is therefore a local
consistency condition on a proposed BPS-string sector, not a construction of
that sector.

\begin{remark}[What the BPS-string statement does and does not construct]
The tension formula is a consequence of the supersymmetry algebra and
tensor-branch normalization.  It does not construct the interacting
tensionless string theory at \(T_q=0\), nor does it define its operator
algebra.  The tensionless limit is part of the fixed-point data whose
existence must be supplied by a construction or by a set of mutually
consistent protected tests.
\end{remark}

\section{Defect Group, Finite Flux, And Polarization}
\label{sec:six-d-defect-group-polarization}

The tensor-branch root lattice is not the whole discrete topological datum of
a \((2,0)\) theory.  Let \(Q_{\mathfrak g}\) be the root lattice of the simply
laced Lie algebra \(\mathfrak g\), with its integral Cartan pairing
\((\ ,\ )\).  The dual lattice is
\[
  P_{\mathfrak g}
  =
  \{\lambda\in Q_{\mathfrak g}\otimes\mathbb R:
  (\lambda,\alpha)\in\mathbb Z\ \text{for every }\alpha\in Q_{\mathfrak g}\},
\]
the weight lattice.  The finite discriminant group
\begin{equation}
  A_{\mathfrak g}=P_{\mathfrak g}/Q_{\mathfrak g}
  \label{eq:six-d-two-zero-defect-group}
\end{equation}
is the finite group detected by the center of the associated simply connected
compact group.  It carries the nondegenerate finite pairing
\begin{equation}
  b_{\mathfrak g}:A_{\mathfrak g}\times A_{\mathfrak g}
  \longrightarrow \mathbb Q/\mathbb Z,
  \qquad
  b_{\mathfrak g}([\lambda],[\mu])=(\lambda,\mu)\bmod\mathbb Z .
  \label{eq:six-d-defect-group-pairing}
\end{equation}
The expression is independent of the representatives because adding a root to
either entry changes \((\lambda,\mu)\) by an integer.  Nondegeneracy is the
statement that \(P_{\mathfrak g}\) is the full dual of \(Q_{\mathfrak g}\).

The ADE values are
\begin{equation}
\begin{array}{c|c}
  \mathfrak g & A_{\mathfrak g}\\ \hline
  A_{N-1} & \mathbb Z_N\\
  D_N,\ N\ \text{odd} & \mathbb Z_4\\
  D_N,\ N\ \text{even} & \mathbb Z_2\oplus\mathbb Z_2\\
  E_6 & \mathbb Z_3\\
  E_7 & \mathbb Z_2\\
  E_8 & 0 .
\end{array}
\label{eq:six-d-ade-defect-groups}
\end{equation}
The order of \(A_{\mathfrak g}\) is \(\det C_{\mathfrak g}\), where
\(C_{\mathfrak g}\) is the Cartan matrix.  Thus the displayed table is the
finite discriminant form of the root lattice, not a new continuous parameter
of the tensor-branch metric.

The group \(A_{\mathfrak g}\) is often described as the group of surface
defect charges modulo dynamical root-lattice strings.  This sentence should
be read as a statement about a protected topological sector: the monograph is
not using it as a construction of the interacting fixed point.  The local
tensor-branch EFT sees strings of charge \(Q_{\mathfrak g}\).  The quotient
\(P_{\mathfrak g}/Q_{\mathfrak g}\) records which finite surface charge can
be measured at infinity after dynamical root-lattice strings have been
allowed to end or screen.

On a closed oriented six-manifold \(Y\), the finite flux sector associated
with this discriminant group is
\begin{equation}
  K_Y(\mathfrak g)=H^3(Y;A_{\mathfrak g}) .
  \label{eq:six-d-finite-flux-group}
\end{equation}
Cup product and \eqref{eq:six-d-defect-group-pairing} define a skew pairing
\begin{equation}
  \langle x,y\rangle_Y
  =
  \int_Y b_{\mathfrak g}(x\smile y)
  \in\mathbb Q/\mathbb Z .
  \label{eq:six-d-finite-flux-pairing}
\end{equation}
The skew sign comes from the odd degree:
\[
  x\smile y=-\,y\smile x
  \qquad\text{for }x,y\in H^3(Y;A_{\mathfrak g}),
\]
while \(b_{\mathfrak g}\) is symmetric.  A self-dual finite flux theory is
not an ordinary sum over all of \(K_Y(\mathfrak g)\).  It determines a finite
Heisenberg representation whose commutator is
\[
  U_xU_yU_x^{-1}U_y^{-1}
  =
  \exp\!\bigl(2\pi\ii\langle x,y\rangle_Y\bigr).
\]
To obtain an ordinary numerical partition function one must supply a
polarization, namely a maximal isotropic subgroup
\[
  L\subset K_Y(\mathfrak g),
  \qquad
  \langle L,L\rangle_Y=0,
  \qquad
  |L|^2=|K_Y(\mathfrak g)|
\]
when the finite pairing on \(K_Y(\mathfrak g)\) is nondegenerate.  Changing
the polarization is a finite global datum, analogous in role to choosing a
global form and discrete theta sector after compactification.  It is not
visible in the local superconformal algebra or in the real tensor-branch
metric alone.

\paragraph{Cyclic finite-flux model.}
For type \(A_{N-1}\), choose a generator of
\(A_{\mathfrak g}\simeq\mathbb Z_N\) and write the finite pairing as
\[
  b(a,b)=\frac{ab}{N}\bmod\mathbb Z
\]
up to the harmless sign determined by the chosen generator.  In the elementary
finite model
\[
  K=\mathbb Z_N\oplus\mathbb Z_N
\]
with coordinates \((a,b)\), the skew pairing is
\begin{equation}
  \langle(a,b),(a',b')\rangle
  =
  \frac{ab'-ba'}{N}\bmod\mathbb Z .
  \label{eq:six-d-cyclic-finite-flux-pairing}
\end{equation}
It is alternating and nondegenerate.  The subgroup
\[
  L=\mathbb Z_N\oplus0
\]
is isotropic and has \(|L|=N\), while \(|K|=N^2\); hence \(L\) is a
polarization.  This finite model is the algebraic core of the polarization
choice.  In an actual six-dimensional compactification the cohomology group
\(H^3(Y;A_{\mathfrak g})\), torsion linking forms, boundary conditions, and
defect insertions must all be specified before the phrase ``the partition
function'' denotes a number rather than a vector in a finite Heisenberg
representation.

\section{Circle Compactification And Five-Dimensional Yang--Mills}

Compactification on a smooth compact manifold \(Y\) gives lower-dimensional
QFTs whose protected data test the six-dimensional proposal.  For a circle
of radius \(R\), a \((2,0)\) theory of type \(\mathfrak g\) yields
five-dimensional maximally supersymmetric Yang--Mills on its Coulomb branch
with gauge algebra \(\mathfrak g\).  This statement is a compactification
test and an effective description, not a Lagrangian definition of the
six-dimensional fixed point.

\paragraph{Five-dimensional coupling convention.}
Use the monograph trace-delta convention for the five-dimensional gauge
action,
\[
  S_{5d}
  =
  \frac1{4g_5^2}
  \int \operatorname{tr}_{\delta}(F_{\mu\nu}F^{\mu\nu})
  +\cdots .
\]
If one unit of five-dimensional instanton number is identified with one unit
of Kaluza--Klein momentum on a circle of radius \(R\), then
\[
  g_5^2=4\pi^2R .
\]
In the common half-trace convention this same relation is
\[
  g_{5,\rm ht}^2=8\pi^2R .
\]

For a static self-dual unit instanton on the spatial \(\mathbb R^4\), the
trace-delta normalization used in the BPST calculation gives
\[
  E_{\rm inst}
  =
  \frac{4\pi^2}{g_5^2}.
\]
Kaluza--Klein momentum on a circle of radius \(R\) is quantized in units
\[
  E_{\rm KK}=\frac1R
\]
for a BPS state with one unit of momentum and no additional transverse
motion.  Identifying the instanton number with KK momentum gives
\[
  \frac{4\pi^2}{g_5^2}=\frac1R,
\]
hence \(g_5^2=4\pi^2R\).  The half-trace convention has
\(g_{5,\rm ht}^2=2g_5^2\), so \(g_{5,\rm ht}^2=8\pi^2R\).

This proportionality is often quoted without specifying the trace convention.
Here it is tied to the instanton normalization used elsewhere in the monograph,
so that
compactification tests, instanton-soliton masses, and Yang--Mills kinetic
terms use the same coupling coordinate.

\paragraph{Wrapped strings and Coulomb-branch masses.}
Compactify a \((2,0)\) theory of type \(\mathfrak g\) on a circle of radius
\(R\), and work on a generic tensor branch.  A BPS string of root charge
\(\alpha\) wrapped once on the circle gives a five-dimensional BPS particle
with mass
\[
  M_\alpha^{\rm wrap}=2\pi R\,|\alpha(\phi_{6d})| .
\]
If this particle is identified with the \(W\)-boson of five-dimensional
maximally supersymmetric Yang--Mills on its Coulomb branch, then the
five-dimensional Coulomb scalar coordinate is normalized by
\[
  \phi_{5d}=2\pi R\,\phi_{6d},
\]
so that \(M_\alpha^{W}=|\alpha(\phi_{5d})|\).  This is a compactification
matching condition; it is not an independent six-dimensional Lagrangian
definition.

A static string of tension \(T_\alpha\) wrapped on a circle of circumference
\(2\pi R\) has rest energy
\[
  M_\alpha^{\rm wrap}=(2\pi R)T_\alpha .
\]
Using Proposition~\ref{prop:six-d-bps-string-tension-central-charge} gives
the first displayed formula.  In a five-dimensional gauge theory on the
Coulomb branch, the mass of the vector multiplet associated with a root
\(\alpha\) is the absolute value of the corresponding root evaluated on the
Coulomb scalar.  Equating the wrapped-string particle with this \(W\)-boson
therefore fixes the scalar normalization
\(\phi_{5d}=2\pi R\,\phi_{6d}\).

\section{Compactification On Riemann Surfaces}

For compactification on a Riemann surface \(C\), partial twists use an
embedding of the spin connection of \(C\) into \(G_R\).  The resulting
four-dimensional theory must have matching anomalies, flavor symmetries,
defect descendants, and moduli-space dimensions.  For \((2,0)\) theories this
is the field-theory origin of class-\(S\) constructions: the input is a
six-dimensional theory of type \(\mathfrak g\), a punctured or unpunctured
Riemann surface, and boundary conditions or defects at punctures.

The compactification is a test with substantial content.  It must specify:
\[
\begin{gathered}
  \text{the twist homomorphism from }\operatorname{Spin}(2)_C\text{ to }G_R,\\
  \text{the defect or puncture data},\\
  \text{the surviving supercharges and flavor symmetries},\\
  \text{the map of six-dimensional anomalies to four-dimensional anomalies},\\
  \text{and the descent of surface defects and BPS strings.}
\end{gathered}
\]
Supplying only a Seiberg--Witten curve in four dimensions does not by itself
construct the six-dimensional parent theory.

\paragraph{Twist datum and anomaly pushforward.}
Here is the local anomaly calculation behind the preceding list in the
unpunctured case.  Let
\[
  M_6=M_4\times C
\]
with \(C\) a closed oriented Riemann surface.  Denote by
\[
  t=e(TC)\in H^2(C;\mathbb Z),
  \qquad
  \int_C t=\chi(C)=2-2g_C,
\]
the Euler class of \(C\).  Decompose the rank-five \(SO(5)_R\) background
bundle as
\[
  N_R=N_2\oplus N_3 ,
\]
where \(N_2\) is an oriented real rank-two bundle for the \(SO(2)_r\)
subgroup and \(N_3\) is an oriented real rank-three bundle for the
four-dimensional \(SO(3)_R\simeq SU(2)_R/\mathbb Z_2\) background.  Write
\[
  x=e(N_2|_{M_4}),\qquad u=p_1(N_3|_{M_4}).
\]
The twist chooses a sign \(\eta=\pm1\), equivalently one of the two
conjugate scalar supercharges after the compactification, and sets the
rank-two Euler class on \(M_4\times C\) to
\begin{equation}
  e(N_2)=x+\eta\,t .
  \label{eq:six-d-class-s-twist-euler-class}
\end{equation}
The sign is a convention tied to the orientation of \(C\) and to the choice
of \(SO(2)_r\) weight.  The formula below keeps \(\eta\) visible so that no
physical conclusion is hidden in an unstated sign convention.

The characteristic classes entering
\eqref{eq:six-d-two-zero-anomaly} are then, modulo terms of degree too high
on \(C\),
\begin{align}
  p_1(N_R)
  &=
  (x+\eta t)^2+u
   =
  x^2+u+2\eta x\,t, \label{eq:six-d-twisted-p1-nr}\\
  p_2(N_R)
  &=
  (x+\eta t)^2u
   =
  x^2u+2\eta x\,u\,t . \label{eq:six-d-twisted-p2-nr}
\end{align}
For the tangent bundle,
\[
  p_1(T(M_4\times C))=p_1(TM_4),
  \qquad
  p_2(T(M_4\times C))=p_2(TM_4),
\]
after integration over \(C\), because \(t^2\) has no component on the
two-dimensional surface.  Let \(p=p_1(TM_4)\).  The coefficient of \(t\) in
the free tensor polynomial \eqref{eq:six-d-free-tensor-anomaly} is
\begin{equation}
  \bigl[I_8(1)\bigr]_{t}
  =
  \frac{\eta x\,t}{48}\,
  \bigl(x^2-p+3u\bigr),
  \label{eq:six-d-free-tensor-twist-linear-term}
\end{equation}
and the coefficient of \(t\) in the interacting excess is
\begin{equation}
  \left[
    \frac{d_{\mathfrak g}h^\vee_{\mathfrak g}}{24}p_2(N_R)
  \right]_{t}
  =
  \frac{\eta d_{\mathfrak g}h^\vee_{\mathfrak g}}{12}\,
  x\,u\,t .
  \label{eq:six-d-excess-twist-linear-term}
\end{equation}
Therefore the six-form anomaly polynomial of the smooth unpunctured
four-dimensional compactification, in this normalization of \(x\) and \(u\),
is
\begin{equation}
  I_6^{\rm class\,S}(\mathfrak g,C;\eta)
  =
  \eta\,\chi(C)\,x
  \left[
    \frac{r_{\mathfrak g}}{48}(x^2-p+3u)
    +
    \frac{d_{\mathfrak g}h^\vee_{\mathfrak g}}{12}u
  \right].
  \label{eq:six-d-class-s-anomaly-pushforward}
\end{equation}
If one rewrites the \(SO(3)_R\) class as
\(u=-4c_2(SU(2)_R)\), this becomes the conventional
four-dimensional \(\mathcal N=2\) anomaly-polynomial coordinate system, up to
the same \(\eta\) and orientation convention.  Punctures, codimension-two
defects, and irregular singularities add localized contributions; they are
not contained in the smooth pushforward \eqref{eq:six-d-class-s-anomaly-pushforward}.

\section{Hitchin-Base Coulomb Operators In Class-\texorpdfstring{\(S\)}{S} Compactification}
\label{sec:six-d-class-s-hitchin-base}

Consider the \((2,0)\) theory of simply laced type \(\mathfrak g\)
compactified on a closed Riemann surface \(C\) of genus \(g_C>1\), with the
standard partial twist that preserves four-dimensional \(\mathcal N=2\)
supersymmetry and with no punctures.  Let
\[
  d_1,\ldots,d_{r_{\mathfrak g}}
\]
be the degrees of the algebraically independent invariant polynomials on
\(\mathfrak g_{\mathbb C}\), ordered increasingly.  Equivalently
\[
  d_i=m_i+1
\]
where \(m_i\) are the exponents of \(\mathfrak g\).  The protected Coulomb
coordinates in the compactified theory are modeled by the Hitchin-base
spaces
\begin{equation}
  \varphi_{d_i}\in H^0(C,K_C^{d_i}),
  \qquad i=1,\ldots,r_{\mathfrak g},
  \label{eq:class-s-hitchin-base-coordinates}
\end{equation}
where \(K_C\) is the holomorphic canonical bundle of \(C\).  In the
four-dimensional theory the coordinate associated with \(\varphi_{d_i}\) has
scaling dimension \(d_i\).

\begin{remark}[Status of the Hitchin-base statement]
\label{rem:hitchin-base-status-six-d}
Equation \eqref{eq:class-s-hitchin-base-coordinates} is used here as a
protected compactification test of the \((2,0)\) theory, not as a definition
of the six-dimensional QFT.  A construction of the compactified theory must
also specify the Hilbert space, local operators, line and surface defects,
metric dependence before twisting, and the treatment of degenerating complex
structures.  The Hitchin base records the Coulomb chiral ring coordinates
visible in the twisted protected sector.
\end{remark}

\paragraph{Riemann--Roch count for Hitchin differentials.}
Let \(C\) be a closed Riemann surface of genus \(g_C>1\).  For every integer
\(d\ge2\),
\begin{equation}
  \dim_{\mathbb C}H^0(C,K_C^d)=(2d-1)(g_C-1).
  \label{eq:hitchin-differential-dimension}
\end{equation}

The degree of \(K_C\) is \(2g_C-2\), hence
\[
  \deg K_C^d=d(2g_C-2).
\]
Riemann--Roch gives
\begin{equation}
  h^0(C,K_C^d)-h^0(C,K_C^{1-d})
  =
  \deg K_C^d+1-g_C .
  \label{eq:rr-hitchin-differential}
\end{equation}
For \(d\ge2\), the line bundle \(K_C^{1-d}\) has negative degree, so it has
no nonzero holomorphic sections.  Therefore
\[
  h^0(C,K_C^d)
  =
  d(2g_C-2)+1-g_C
  =
  (2d-1)(g_C-1).
\]

\paragraph{Degree sum for a simple Lie algebra.}
For a complex simple Lie algebra \(\mathfrak g_{\mathbb C}\) with invariant
polynomial degrees \(d_i=m_i+1\),
\begin{equation}
  \sum_{i=1}^{r_{\mathfrak g}}(2d_i-1)=d_{\mathfrak g},
  \label{eq:lie-degree-sum-dimension}
\end{equation}
where \(d_{\mathfrak g}=\dim_{\mathbb C}\mathfrak g_{\mathbb C}\).

The exponents \(m_i\) satisfy
\begin{equation}
  \sum_{i=1}^{r_{\mathfrak g}}m_i=|\Delta_+|,
  \label{eq:exponent-sum-positive-roots}
\end{equation}
where \(\Delta_+\) is a set of positive roots.  One way to see
\eqref{eq:exponent-sum-positive-roots} is to compare the top degree in the
standard Poincare polynomial identity for the flag variety:
\[
  \sum_{w\in W_{\mathfrak g}}t^{2\ell(w)}
  =
  \prod_{i=1}^{r_{\mathfrak g}}
  \frac{1-t^{2(m_i+1)}}{1-t^2}.
\]
The maximal length of a Weyl-group element is \(|\Delta_+|\), while the top
degree on the product side is \(2\sum_i m_i\).  Hence
\(\sum_i m_i=|\Delta_+|\).  Since
\[
  d_{\mathfrak g}=r_{\mathfrak g}+2|\Delta_+|,
\]
we obtain
\[
  \sum_i(2d_i-1)
  =
  \sum_i(2m_i+1)
  =
  2|\Delta_+|+r_{\mathfrak g}
  =
  d_{\mathfrak g}.
\]

\paragraph{Unpunctured class-\(S\) Coulomb-base dimension.}
For the unpunctured genus-\(g_C>1\) compactification above, the protected
Coulomb base modeled by the Hitchin base
\begin{equation}
  \mathcal B_{\mathfrak g,C}
  =
  \bigoplus_{i=1}^{r_{\mathfrak g}} H^0(C,K_C^{d_i})
  \label{eq:class-s-hitchin-base}
\end{equation}
has complex dimension
\begin{equation}
  \dim_{\mathbb C}\mathcal B_{\mathfrak g,C}
  =
  (g_C-1)d_{\mathfrak g}.
  \label{eq:class-s-coulomb-dimension}
\end{equation}

By the Riemann--Roch calculation above,
\begin{equation}
  \dim_{\mathbb C}\mathcal B_{\mathfrak g,C}
  =
  \sum_i(2d_i-1)(g_C-1).
  \label{eq:hitchin-base-dimension-sum}
\end{equation}
Factoring out \(g_C-1\) and using the degree-sum calculation above gives
\[
  \dim_{\mathbb C}\mathcal B_{\mathfrak g,C}
  =
  (g_C-1)\sum_i(2d_i-1)
  =
  (g_C-1)d_{\mathfrak g}.
\]
The assumptions \(g_C>1\) and \(d_i\ge2\) are used in the Riemann--Roch
lemma to remove the \(H^0(C,K_C^{1-d_i})\) term; punctures or genus zero/one
surfaces require additional divisor data and obey different dimension
formulae.

\begin{example}[ADE degree checks]
\label{ex:ade-class-s-coulomb-dimension-checks}
For \(A_{N-1}\), the degrees are \(2,3,\ldots,N\).  Therefore
\[
  \sum_{d=2}^{N}(2d-1)=N^2-1=d_{A_{N-1}},
\]
and the unpunctured genus-\(g_C\) Coulomb-base dimension is
\[
  (g_C-1)(N^2-1).
\]
For \(D_N\), the degrees are
\[
  2,4,6,\ldots,2N-2,\ N,
\]
with the usual repeated degree when \(N\) is even.  Their contribution is
\[
  \sum_{j=1}^{N-1}(4j-1)+(2N-1)
  =
  N(2N-1)=d_{D_N}.
\]
For the exceptional algebras, the degree lists are
\[
\begin{array}{c|c}
  \mathfrak g & (d_i)\\ \hline
  E_6 & 2,5,6,8,9,12\\
  E_7 & 2,6,8,10,12,14,18\\
  E_8 & 2,8,12,14,18,20,24,30 .
\end{array}
\]
In each case \(\sum_i(2d_i-1)\) gives \(78\), \(133\), and \(248\),
respectively.
\end{example}

\section{Status Ledger}

Six-dimensional SCFT data in this monograph consist of:
\[
\begin{gathered}
  \text{superconformal representation category},\\
  \text{tensor-branch EFT with self-dual fields and Green--Schwarz terms},\\
  \text{anomaly polynomial and flavor-background data},\\
  \text{BPS string charge lattice and inflow},\\
  \text{finite defect group, flux pairing, and polarization data},\\
  \text{compactification tests and extended-operator descendants}.
\end{gathered}
\]
For \((2,0)\) type \(\mathfrak g\) this includes the simply laced Lie algebra,
the tensor-branch quotient
\((\mathbb R^5\otimes\mathfrak t_{\mathfrak g})/W_{\mathfrak g}\), the
anomaly polynomial \eqref{eq:six-d-two-zero-anomaly}, the root-lattice BPS
strings, the discriminant group
\(A_{\mathfrak g}=P_{\mathfrak g}/Q_{\mathfrak g}\) with finite pairing
\eqref{eq:six-d-defect-group-pairing}, and the five-dimensional
compactification normalization \(g_5^2=4\pi^2R\) in trace-delta variables.

No item in this list is dispensable.  A proposed fixed point that supplies
only a branch EFT is incomplete; a proposed non-Lagrangian object that
supplies only protected indices is also incomplete.

\begin{openproblem}[Six-dimensional construction]
\label{op:six-dimensional-construction}
Construct an interacting six-dimensional SCFT as a local QFT with Hilbert
space, local operator algebra, stress tensor, anomaly polynomial, tensor
branch, BPS strings, and compactification functoriality in one framework.
\end{openproblem}
