\chapter{Moduli Spaces In Supersymmetric Quantum Field Theory}
\label{ch:susy-moduli-spaces}

\ConventionsMixed

This chapter defines supersymmetric moduli spaces as spaces of supersymmetric
vacua together with their low-energy QFT structures.  The set of classical solutions
is only the first coordinate chart; the physical object includes quotienting,
singular strata, massless fields, chiral rings, metrics, and quantum
corrections.

\begin{definition}[Quantum supersymmetric moduli-space structure]
\label{def:quantum-supersymmetric-moduli-space-datum}
Fix a supersymmetric local QFT with supercharges \(Q_\alpha\) acting on a
common invariant domain in each vacuum representation.  A quantum
supersymmetric moduli-space structure is a tuple
\[
  \mathfrak M
  =
  \bigl(
  \mathcal M,\mathcal O_{\mathcal M}^{\mathrm{hol}},
  \{\mathcal H_p,\mathcal A_p,\mathcal E_p,\mathcal G_p\}_{p\in\mathcal M}
  \bigr)
\]
with the following entries.
\begin{enumerate}
\item \(\mathcal M\) is a reduced stratified analytic space whose points
label equivalence classes of supersymmetric vacuum sectors.  A representative
vacuum sector is a translation-invariant positive-energy representation with
vacuum vector or vacuum state annihilated by the preserved supercharges,
modulo unitary equivalence preserving the local operator net and background
couplings.
\item \(\mathcal O_{\mathcal M}^{\mathrm{hol}}\) is the sheaf of holomorphic
branch-coordinate functions generated, where valid, by expectation values of
protected chiral operators.  Nilpotent or derived structure is additional
structure; if
it affects operator products or low-energy dynamics it must be retained
rather than replaced by the reduced space.
\item \(\mathcal H_p\) is the Hilbert space of fluctuations in the vacuum
sector \(p\), and \(\mathcal A_p\) is the local operator algebra or field
presentation of the low-energy QFT in that sector.
\item \(\mathcal E_p\) records the effective branch structure: massless fields,
two-derivative metric, Wess--Zumino terms, superselection sectors, and the
energy range on which the branch effective theory is valid.
\item \(\mathcal G_p\) records background responses, anomaly lines, and
global-form or extended-operator structures in the vacuum sector.
\end{enumerate}
On a smooth stratum, \(T_p\mathcal M\) is represented by supersymmetric
massless scalar directions in \(\mathcal H_p\).  A singular stratum is a locus
where this representation, the effective theory \(\mathcal E_p\), or the
background-response structure \(\mathcal G_p\) changes.  A classical quotient of
scalar zero modes is therefore a coordinate construction for one approximation
to \(\mathfrak M\), not the definition of the quantum moduli-space structure.
\end{definition}

\section{Vacuum Equations}

For four-dimensional \(\mathcal N=1\) chiral multiplets \(\Phi^i\) with
superpotential \(W(\Phi)\) and gauge group \(G\), a classical supersymmetric
vacuum obeys
\[
  F_i=\frac{\partial W}{\partial\Phi^i}=0,
  \qquad
  \mu^a=0,
\]
where \(\mu:\mathcal X\to\mathfrak g^\ast\) is the moment map for the
K\"ahler target \(\mathcal X\) of scalar fields.  The classical vacuum space is
the K\"ahler quotient
\[
  \mathcal M_{\mathrm{cl}}
  =
  \{F=0,\ \mu=0\}/G .
\]
Equivalently, under the required stability hypotheses for the complexified gauge
action, it is the holomorphic quotient
\[
  \{F=0\}/G_\mathbb C .
\]
The second description identifies holomorphic functions on
\(\mathcal M_{\mathrm{cl}}\) with gauge-invariant chiral polynomials modulo
the ideal generated by the \(F\)-term equations.

The last sentence hides several hypotheses, so we spell out the quotient
construction used in this volume.  Let \(V\) be the finite-dimensional Hermitian
representation of \(G\) carried by the scalar fields, let \(G_\mathbb C\) be a
reductive complexification, and let \(W\in\mathbb C[V]^{G_\mathbb C}\) be a
polynomial superpotential.  Write
\[
  Z_F=\{v\in V\mid dW(v)=0\},\qquad
  I_F=(\partial_1W,\ldots,\partial_nW)\subset\mathbb C[V].
\]
The affine chiral quotient is
\begin{equation}
  \mathcal M_{\mathrm{ch},\mathrm{cl}}
  =
  \operatorname{Spec}
  \left((\mathbb C[V]/I_F)^{G_\mathbb C}\right)_{\mathrm{red}},
  \label{eq:affine-chiral-quotient}
\end{equation}
where the subscript records that the displayed affine variety forgets
nilpotents unless they are kept as derived or scheme-theoretic data.  The
symplectic vacuum quotient is
\begin{equation}
  \mathcal M_{\mu,\mathrm{cl}}
  =
  (Z_F\cap\mu^{-1}(0))/G .
  \label{eq:symplectic-vacuum-quotient}
\end{equation}

\paragraph{Classical quotient and chiral functions.}
Assume the Kempf--Ness comparison for the triple
\((V,G,G_\mathbb C)\) on the \(G_\mathbb C\)-invariant closed subset
\(Z_F\): every polystable \(G_\mathbb C\)-orbit in \(Z_F\) meets
\(\mu^{-1}(0)\), and two points of \(Z_F\cap\mu^{-1}(0)\) lie in the same
closed \(G_\mathbb C\)-orbit exactly when they lie in the same \(G\)-orbit.
Under this comparison \(\mathcal M_{\mu,\mathrm{cl}}\) is naturally identified with the
polystable affine quotient \(\mathcal M_{\mathrm{ch},\mathrm{cl}}\), and its
holomorphic coordinate ring is
\[
  \mathbb C[\mathcal M_{\mathrm{ch},\mathrm{cl}}]
  =
  \left((\mathbb C[V]/I_F)^{G_\mathbb C}\right)_{\mathrm{red}} .
\]
Thus the notation \(\{F=0\}/G_\mathbb C\) in this monograph means the affine
quotient in \eqref{eq:affine-chiral-quotient}, together with the stated
stability hypothesis, not the naive set of all complexified orbits.
Because \(W\) is \(G_\mathbb C\)-invariant, the differential \(dW\) is
equivariant and the ideal \(I_F\) is \(G_\mathbb C\)-stable.  Hence every
class in \((\mathbb C[V]/I_F)^{G_\mathbb C}\) is represented by a
gauge-invariant holomorphic polynomial on \(Z_F\), and restriction to
\(Z_F\cap\mu^{-1}(0)\) gives a holomorphic function constant on \(G\)-orbits.
The assumed Kempf--Ness statement identifies \(G\)-orbits in
\(Z_F\cap\mu^{-1}(0)\) with closed \(G_\mathbb C\)-orbits in the affine
quotient.  Invariant polynomials separate closed orbits in an affine
reductive quotient, so the descended functions are precisely the regular
holomorphic functions on the reduced quotient.  Nonclosed complexified orbits
can have the same closure and therefore the same values of all invariant
polynomials; the affine quotient collapses them to the polystable
representative.

\begin{example}[Rank-one abelian quotient]
\label{ex:susy-rank-one-abelian-quotient}
Let \(G=U(1)\) act on \(V=\mathbb C^2\) by
\[
  (x,y)\longmapsto(\ee^{\ii\alpha}x,\ee^{-\ii\alpha}y),
  \qquad W=0,
\]
with moment map \(\mu=|x|^2-|y|^2\).  The invariant ring of the
complexified action \(t\cdot(x,y)=(tx,t^{-1}y)\) is
\[
  \mathbb C[x,y]^{\mathbb C^\ast}=\mathbb C[u],
  \qquad u=xy .
\]
Indeed the monomial \(x^ay^b\) has \(U(1)\) charge \(a-b\), so invariance
forces \(a=b\) and the monomial is \(u^a\).  On the symplectic side
\(\mu=0\) gives \(|x|=|y|\).  The map \((x,y)\mapsto xy\) sends
\(\mu^{-1}(0)/U(1)\) bijectively to \(\mathbb C\): for
\(u=\rho\ee^{\ii\theta}\), with \(\rho\ge0\), choose
\[
  x=\sqrt{\rho},\qquad y=\sqrt{\rho}\,\ee^{\ii\theta},
\]
and changing the phase split between \(x\) and \(y\) is precisely the
\(U(1)\) quotient.  Thus this basic vacuum quotient is the complex line with
coordinate \(u\), not a two-complex-dimensional space of unquotiented scalar
expectation values.
\end{example}

\section{Quantum Moduli Space}

For the quantum moduli-space structure
\(\mathfrak M\) of
Definition~\ref{def:quantum-supersymmetric-moduli-space-datum}, a smooth
branch is a stratum on which the massless field content, preserved
supersymmetry algebra, and background-response structure are locally constant.
If \(z^i\) are massless scalar coordinates on such a branch, the
two-derivative effective action determines a metric
\[
  ds^2=g_{i\bar j}(z,\bar z)\,dz^i d\bar z^{\bar j}
\]
on the branch.  The metric is part of the low-energy QFT structure; the chiral
coordinate ring determines only holomorphic functions.

\section{Chiral Rings And Holomorphic Functions}

Let \(\mathcal R_{\mathrm{ch}}\) be the chiral ring of gauge-invariant local
operators modulo \(\bar Q\)-exact operators.  In a vacuum \(p\), expectation
values define a homomorphism
\[
  \operatorname{ev}_p:\mathcal R_{\mathrm{ch}}\longrightarrow\mathbb C .
\]
If chiral operators separate vacua on a branch and have no nilpotent
relations in the reduced branch coordinate ring, the collection of
\(\operatorname{ev}_p\) identifies the reduced holomorphic coordinate ring
with a quotient of \(\mathcal R_{\mathrm{ch}}\).  Nilpotents, accidental
massless fields, and singular strata must be retained when they affect
operator products or low-energy dynamics.

\section{Fixed Charge As A Probe Of A Branch}

There are two distinct operator-theoretic probes of a moduli branch.  The
first probe is the character of the chiral ring in a vacuum,
\(\operatorname{ev}_p:\mathcal R_{\mathrm{ch}}\to\mathbb C\).  It addresses
the reduced holomorphic coordinate algebra, under the separation hypotheses
stated above.  The second probe uses sectors of fixed global charge.  This
second probe is the precise setting of the large-charge expansion: a
low-energy effective theory on a smooth branch can determine the asymptotic
spectrum of highly charged states, and in a conformal theory those states
are equivalently highly charged local operators of definite scaling
dimension.

Let \(F\) be a compact internal global symmetry of the QFT.  A fixed charge
sector is not defined by diagonalizing all of \(\mathfrak f=\operatorname{Lie}F\).
One first chooses a torus \(T\subset F\), usually a maximal torus, and the
global form of \(F\) determines the allowed weight lattice
\[
  \Lambda_{\rm QFT}\subset \Lambda_T=\operatorname{Hom}(T,U(1)).
\]
For \(F=\widetilde F/Z\), for example, only the \(\widetilde F\)-weights
that are trivial on \(Z\) are charges of honest \(F\)-representations.
Projective actions, mixed anomalies, background fluxes, and discrete theta
angles can shift this lattice or replace it by twisted sectors; that data is
part of the sector definition, not a correction to be added after the saddle
has been written.  For nonabelian \(F\), the invariant label of a charge
sector is a Weyl orbit \([q]\in\Lambda_{\rm QFT}/W_F\), often represented by
a dominant weight \(q\).

Choose an integral basis \(e_A\) of the selected torus and define
\[
  Q_A=\int_{\Sigma} d\Sigma_\mu\, j_A^\mu ,
  \qquad A=1,\ldots,r ,
\]
on a spatial slice \(\Sigma\), after specifying the domain, regulator, and
background bundle.  For \(q\in\Lambda_{\rm QFT}\) write
\[
  \Hilb_q=\{\psi\in\Hilb\mid Q_A\psi=q_A\psi\ \text{for all }A\}.
\]
The Hamiltonian chemical potentials \(\mu^A\) are real Lagrange multipliers
in
\[
  H_\mu=H-\mu^A Q_A .
\]
In an Euclidean finite-volume path integral the same source is continued to
a flat background holonomy, \(A_\tau=i\mu^A e_A\) in the simplest
reflection-positive convention.  Wess-Zumino terms and Chern-Simons/contact
terms can add background-charge contributions, so the Euclidean holonomy
ledger must be matched to the Lorentzian Noether normalization before one
compares \(q_A\) with a lattice vector.

Assume now that a smooth branch \(\mathcal M_{\rm sm}\subset\mathcal M\)
carries the chosen \(T\)-action, generated by real Killing fields
\[
  K_A=K_A^i(z)\partial_i+\overline{K}_A^{\bar i}(\bar z)\partial_{\bar i},
\]
normalized so that \(\exp(2\pi e_A)\) is the identity on fields of unit
charge.  Let the branch EFT with metric \(g_{i\bar j}\), possible
Wess-Zumino terms, and background responses be valid below the transverse
mass scale \(M_{\rm gap}(p)\) at \(p\in\mathcal M_{\rm sm}\).  A homogeneous
charged ansatz on \(\mathbb R\times S_R^{d-1}\) has
\[
  \dot z^i=\mu^A K_A^i(z),\qquad
  \dot{\bar z}^{\bar i}=\mu^A\overline{K}_A^{\bar i}(\bar z),
\]
only after the auxiliary-field equations, preserved-supercharge equations,
and possible Gauss-law constraints have been imposed.  If an emergent gauge
field remains massless on the branch, the fixed-charge problem is solved on
the gauge-invariant phase space: electric flux sectors, moment-map
constraints, and boundary charges must be fixed together with the global
charge.

At the two-derivative level, with \(V_R=\operatorname{Vol}(S_R^{d-1})\), the
Noether map has the following convention-dependent form
\begin{align}
  q_A
  &=
  V_R\,g_{i\bar j}(z,\bar z)
  \left(K_A^i\dot{\bar z}^{\bar j}
        +\dot z^i\overline{K}_A^{\bar j}\right)
  +q_A^{\rm WZ}+q_A^{\rm bg}
  +\cdots ,
  \label{eq:moduli-large-charge-noether-map}\\
  E_{\rm EFT}(q,R)
  &=
  \underset{z,\mu}{\operatorname{crit}}
  \left[
    \mu^A q_A
    -L_{\rm br}(z,\mu;R)
  \right] .
  \label{eq:moduli-large-charge-epsilon}
\end{align}
Here \(L_{\rm br}(z,\mu;R)\) is the branch Lagrangian evaluated on the
homogeneous orbit, including curvature couplings, Wess-Zumino terms,
background responses, and higher-derivative operators.  The saddle is
locally stable only when the susceptibility matrix
\(\partial q_A/\partial\mu^B\) on the constrained branch has the correct
positive signature and the quadratic fluctuation operator has no negative
modes after the collective coordinates are separated.  If several Weyl
images, disconnected components, or symmetry-breaking patterns solve the
same fixed-charge problem, \(E_0(q,R)\) is the lowest of the corresponding
Routhians; the other saddles describe excited sectors or metastable
large-charge states.

Thus the formal homogeneous saddle is not yet a large-charge expansion.  In
the non-BPS Goldstone class one must exhibit a simultaneous hierarchy
\begin{equation}
  \label{eq:moduli-large-charge-scale-window}
\begin{gathered}
  R^{-1}\ll |\mu|\ll M_{\rm gap}(p(\mu)),
  \\
  \frac{1}{|\mu|R}\ll1,\qquad
  \frac{|\mu|}{M_{\rm gap}(p(\mu))}\ll1 .
\end{gathered}
\end{equation}
For instance, if along the saddle family
\[
  M_{\rm gap}(p(\mu))\ge C|\mu|^{1+\delta}M_0^{-\delta}
  \qquad (C>0,\ \delta>0)
\]
for \(|\mu|R\to\infty\), then both small parameters in
\eqref{eq:moduli-large-charge-scale-window} decrease.  If instead
\(M_{\rm gap}\) is constant, or grows no faster than \(|\mu|\), the
transverse modes enter before the candidate large-charge expansion becomes
parametric.  When \(|\mu|\) reaches such a threshold, those fields must be
included in the low-energy theory.  When a different saddle has lower energy
in the same Weyl orbit \([q]\), the chosen branch no longer computes the
lowest state in \(\Hilb_q\).

A simple supersymmetric branch model shows the distinction between charge
sector, Noether map, and scale window.  Take four-dimensional
\(\mathcal N=1\) chiral fields \(X,Y\) with canonical K\"ahler metric and
\[
  W=\frac{\lambda}{2}XY^2 .
\]
The \(U(1)\) action \(X\mapsto e^{in\alpha}X\), \(Y\mapsto
e^{-in\alpha/2}Y\) is integral for \(n=2\), so \(X\) has charge \(2\) and
\(Y\) has charge \(-1\).  The branch \(Y=0\) is parametrized by \(X\), and
the transverse chiral field has mass
\[
  M_Y(X)=|\lambda X| .
\]
On \(\mathbb R\times S_R^3\), keep the conformal curvature term for the
branch scalar and write \(V_R=2\pi^2R^3\).  For
\[
  X(t)=\rho e^{i\omega t},\qquad \omega=n\mu ,
\]
the branch Lagrangian is
\[
  L_X=V_R\left(\omega^2-\frac1{R^2}\right)\rho^2 ,
\]
and the charge and energy are
\[
  q=2n\omega V_R\rho^2=2n^2\mu V_R\rho^2,\qquad
  E_X=V_R\left(\omega^2+\frac1{R^2}\right)\rho^2 .
\]
At fixed positive branch charge \(q=nJ\), minimization gives
\begin{equation}
  \label{eq:moduli-large-charge-susy-branch-example}
\begin{gathered}
  \omega=\frac1R,\qquad
  \rho^2=\frac{qR}{2nV_R}
        =\frac{q}{4\pi^2nR^2},\qquad
  E_X(q,R)=\frac{q}{nR}=\frac{J}{R}.
\end{gathered}
\end{equation}
The \(X\)-branch oscillators have the conformal scalar spectrum
\(\Omega_{X,\ell}=(\ell+1)/R\) on \(S^3\), with the homogeneous phase mode
treated as the fixed-charge collective coordinate.  The transverse
\(Y\)-multiplet has
\[
  \Omega_{Y,\ell}^2
  =
  \frac{(\ell+1)^2}{R^2}
  +|\lambda|^2\rho^2
  +O(\mu/R,\mu^2),
\]
where the displayed \(O(\mu)\) terms are charge-frame shifts bounded by the
chemical potential.  The branch window is therefore
\[
  M_YR
  =
  |\lambda|\sqrt{\frac{q}{4\pi^2 n}}\gg1,\qquad
  \frac{|\mu|}{M_Y}
  =
  \frac{2\pi}{|\lambda|\sqrt{nq}}\ll1 .
\]
This is a controlled fixed-charge supersymmetric branch saddle for
\(q\gg4\pi^2 n/|\lambda|^2\), but it is not an example of the non-BPS
large-\(\mu\) Goldstone expansion because \(\mu R=1/n\).  The physics lesson
is the point needed for branch reconstruction: one must state which
hierarchy is actually present.  A protected chiral or BPS large-charge
sector may be controlled by \(M_{\rm gap}R\to\infty\) and charge
quantization, while a generic non-BPS large-charge EFT needs the stronger
window \eqref{eq:moduli-large-charge-scale-window}.

In a unitary CFT, put the theory on
\(\mathbb R\times S^{d-1}_R\).  The state-operator theorem of
Chapter~\ref{ch:volume-iii-radial-quantization} identifies local
operators of definite scaling dimension with the corresponding energy
eigenstates in the local-operator subspace of the cylinder Hilbert space.
Consequently, if the lowest-energy vector in \(\Hilb_q\) is represented by
a primary operator \(\mathcal O_q\), then
\[
  \Delta(\mathcal O_q)=R\,E_0(q,R).
\]
The large-charge program initiated by Hellerman and collaborators is exactly
this bridge in the conformal setting: fixed-charge spectral data determine
the coefficients of the branch EFT that control the charged saddle, while
the branch EFT predicts the asymptotic dimensions and matrix elements of
large-charge operators.  Its output is an asymptotic charged-sector
reconstruction.  Recovering the full vacuum datum
\((\mathcal M,\mathcal H_p,\mathcal A_p,\mathcal G_p)\), including
singular strata, uncharged directions, nilpotent scheme structure,
background responses, and changes of low-energy degrees of freedom, requires
additional input beyond the large-charge sector alone.

The broader reconstruction problem is more structural.  Wightman-type data
provide a Hilbert space, local operator-valued distributions, vacuum
correlation distributions, spectrum, locality, and covariance.  A
Kontsevich--Segal-type theory asks instead for functorial amplitudes assigned
to geometric bordisms, with boundary state spaces, sewing, reflection
positivity, and background-field dependence.  A supersymmetric moduli space
should be visible in such KS data as a family of boundary conditions or
asymptotic sectors whose sewn amplitudes produce the low-energy branch
theory.  Chiral-ring characters and fixed-charge spectra are therefore
probes of the Wightman-to-KS reconstruction map, rather than the map itself.

\section{N One SQCD Branches}

For \(SU(N_c)\) SQCD with \(N_f\) flavors, the microscopic chiral fields are
quarks \(Q^a{}_i\) in the fundamental and antiquarks
\(\widetilde Q_a{}^i\) in the antifundamental, with color
\(a=1,\ldots,N_c\) and flavor \(i=1,\ldots,N_f\).  The elementary
gauge-invariant chiral coordinates are the mesons
\[
  M^i{}_j=\widetilde Q_a{}^i Q^a{}_j
\]
and, when \(N_f\geq N_c\), the baryons and antibaryons obtained by
contracting \(N_c\) quarks or antiquarks with \(\epsilon_{a_1\cdots a_{N_c}}\).
The relations among these coordinates are algebraic relations in the reduced
affine quotient; they are not optional decoration, because they decide which
operator expectation values can occur in a supersymmetric vacuum.

The following table is only an orientation map.  It is not used as a proof,
and each row requires a separate derivation or an explicitly labeled
Wilsonian input:
\[
\begin{array}{c|c}
N_f<N_c & \text{ADS superpotential produces a runaway direction}\\
N_f=N_c & \det M-B\widetilde B=\Lambda_h^{2N_c}\\
N_f=N_c+1 & \text{confining chiral coordinates with superpotential constraint}\\
\end{array}
\]
Each entry is a Wilsonian chiral-coordinate statement.  The metric on the
branch and the massive spectrum require additional low-energy data.

The table becomes meaningful only after the classical quotient coordinates
have been fixed and after the status of the quantum input has been separated
from the algebra done with that input.  The smallest nontrivial
\(N_f=N_c\) example is therefore a concrete test case for the monograph's
method: construct the
classical quotient from invariant theory, state the protected quantum
deformation as the only dynamical input, and then derive all consequences
that follow from the deformed equation by ordinary algebra.

\paragraph{Classical \(SU(2)\), \(N_f=2\) quotient as invariant theory.}
Consider four-dimensional \(\mathcal N=1\) \(SU(2)\) SQCD with two flavors
and \(W=0\).  Package the two quark and two antiquark doublets, using the
pseudoreality of the \(SU(2)\) fundamental, as four chiral doublets
\[
  P_a^I,\qquad a=1,2,\qquad I=1,\ldots,4 .
\]
The complexified gauge group \(SL(2,\mathbb C)\) acts on the gauge index
\(a\).  Define the antisymmetric gauge-invariant coordinates
\begin{equation}
  V^{IJ}=\epsilon^{ab}P_a^I P_b^J=-V^{JI},
  \qquad
  1\leq I,J\leq4,
  \label{eq:su2-nf2-meson-baryon-v-coordinates}
\end{equation}
with \(\epsilon^{12}=1\).  Let
\begin{equation}
  \operatorname{Pf}(V)
  =
  V^{12}V^{34}-V^{13}V^{24}+V^{14}V^{23}.
  \label{eq:su2-nf2-pfaffian-convention}
\end{equation}
Then the reduced affine classical chiral quotient is
\begin{equation}
  \mathcal M_{\mathrm{ch},\mathrm{cl}}
  \simeq
  \{\,V\in\bigwedge\nolimits^2\mathbb C^4
    \mid \operatorname{Pf}(V)=0\,\}.
  \label{eq:su2-nf2-classical-plucker-hypersurface}
\end{equation}
It has complex dimension \(5\).  In the usual \(Q,\widetilde Q\) flavor
splitting, the six entries of \(V^{IJ}\) are the mesons and baryons, and
\eqref{eq:su2-nf2-classical-plucker-hypersurface} is the classical
\(\det M-B\widetilde B=0\) relation in an \(SU(4)\)-covariant notation.
The quantum deformation \(\operatorname{Pf}(V)=\Lambda_h^4\) is not part of
this classical quotient; it is the Wilsonian nonperturbative input discussed
in Chapter~\ref{ch:susy-four-dimensional-n1-gauge-dynamics}.

Write the two rows of the \(2\times4\) matrix \(P\) as
\[
  u=(P_1^1,\ldots,P_1^4),\qquad
  v=(P_2^1,\ldots,P_2^4).
\]
Then \(V^{IJ}=u_Iv_J-u_Jv_I\), so \(V=u\wedge v\) is a decomposable two-form.
Hence \(V\wedge V=0\).  With the convention
\eqref{eq:su2-nf2-pfaffian-convention}, the coefficient of
\(2e_1\wedge e_2\wedge e_3\wedge e_4\) in \(V\wedge V\) is
\(\operatorname{Pf}(V)\), proving that every gauge orbit maps into the
displayed hypersurface.

Conversely, suppose \(V\neq0\) and \(\operatorname{Pf}(V)=0\).  After
relabeling the four flavor indices, assume \(a_0=V^{12}\neq0\).  Set
\[
  u=(1,0,-V^{23}/a_0,-V^{24}/a_0),
  \qquad
  v=(0,a_0,V^{13},V^{14}).
\]
The minors \(u_Iv_J-u_Jv_I\) reproduce \(V^{12}\), \(V^{13}\), \(V^{14}\),
\(V^{23}\), and \(V^{24}\) by construction, while the remaining minor is
\[
  u_3v_4-u_4v_3
  =
  \frac{V^{13}V^{24}-V^{14}V^{23}}{a_0}
  =
  V^{34},
\]
where the last equality is exactly \(\operatorname{Pf}(V)=0\).  The point
\(V=0\) is obtained from \(P=0\).  Thus the invariant map has precisely the
hypersurface \eqref{eq:su2-nf2-classical-plucker-hypersurface} as its image.

It remains to justify that no additional reduced affine-coordinate data are
missing.  In a polynomial invariant, all \(SL(2,\mathbb C)\) gauge indices
must be contracted by invariant tensors.  For the fundamental two-dimensional
representation, the invariant tensors are generated by the alternating tensor
\(\epsilon^{ab}\); equivalently, after decomposing tensor powers into
irreducible \(SL(2)\) representations, the singlet summands are obtained by
pairwise antisymmetric contractions.  Hence the reduced invariant ring is
generated by the pairings \(V^{IJ}\).  For four flavor labels the only
quadratic straightening relation among these six generators is the Plucker
relation
\[
  V^{12}V^{34}-V^{13}V^{24}+V^{14}V^{23}=0,
\]
which we already identified as the Pfaffian equation.  Therefore the reduced
affine quotient is the displayed hypersurface.  Its dimension is
\(6-1=5\), agreeing with the quotient count
\[
  \dim_{\mathbb C}\mathbb C^{8}-\dim_{\mathbb C}SL(2,\mathbb C)
  =
  8-3=5
\]
on the stable maximal-rank locus.

\paragraph{Algebra after the \(SU(2)\), \(N_f=2\) quantum-deformation input.}
Assume the Wilsonian chiral-coordinate input that the quantum branch of the
theory whose classical chiral quotient is
\eqref{eq:su2-nf2-classical-plucker-hypersurface} is
\begin{equation}
  \operatorname{Pf}(V)=\Lambda_h^4,
  \qquad \Lambda_h\neq0,
  \label{eq:su2-nf2-quantum-deformed-pfaffian}
\end{equation}
where \(\operatorname{Pf}\) is the polynomial in
\eqref{eq:su2-nf2-pfaffian-convention}.  The dynamical content is the stated
Wilsonian input; the following paragraphs record only the algebraic
consequences of that input.  The deformed branch is
smooth and five-complex-dimensional.  Moreover, after adding a nondegenerate
antisymmetric mass source and putting it, by a complex linear change of
flavor basis, in Darboux form
\begin{equation}
  W_m=m_1V^{12}+m_2V^{34},
  \qquad m_1m_2\neq0,
  \label{eq:su2-nf2-diagonal-mass-source}
\end{equation}
and imposing the constraint by a chiral Lagrange multiplier \(X\),
\begin{equation}
  W_{\mathrm{constr}}
  =
  m_1V^{12}+m_2V^{34}
  +X\bigl(\operatorname{Pf}(V)-\Lambda_h^4\bigr),
  \label{eq:su2-nf2-constrained-mass-superpotential}
\end{equation}
the \(F\)-term equations over \(\mathbb C\) have exactly two solutions.
After choosing a square root \(\Lambda_h^2\) of the deformation parameter
\(\Lambda_h^4\) and a square root of \(m_1m_2\), they are
\begin{equation}
  X_\pm=\pm\frac{\sqrt{m_1m_2}}{\Lambda_h^2},
  \qquad
  V^{12}_\pm=-\frac{m_2}{X_\pm},
  \qquad
  V^{34}_\pm=-\frac{m_1}{X_\pm},
  \label{eq:su2-nf2-massive-vacua}
\end{equation}
with all other \(V^{IJ}\) equal to zero.  Their superpotential values are
\begin{equation}
  W_\pm
  =
  \mp\,2\Lambda_h^2\sqrt{m_1m_2}.
  \label{eq:su2-nf2-massive-superpotential-values}
\end{equation}
Thus the deformed branch removes the classical origin, and the simplest
mass deformation has the two isolated chiral vacua expected after decoupling
to pure \(SU(2)\) supersymmetric Yang--Mills, with scale matching
\(\Lambda_0^6=m_1m_2\Lambda_h^4\) in this Darboux convention.  The last
comparison uses matching to the massive low-energy theory; the algebraic
statement above uses only
\eqref{eq:su2-nf2-quantum-deformed-pfaffian}.  Different choices of square
root or of the sign convention relating \(V^{IJ}\) to the usual meson-baryon
coordinates only relabel the two vacua.

Let
\[
  F(V)=V^{12}V^{34}-V^{13}V^{24}+V^{14}V^{23}.
\]
The gradient of \(F\) in the six affine coordinates is
\[
  \nabla F
  =
  (V^{34},-V^{24},V^{23},-V^{13},V^{14},V^{12}),
\]
ordered as
\[
  (V^{12},V^{13},V^{14},V^{24},V^{23},V^{34}).
\]
If \(\nabla F=0\), then all six entries of \(V\) vanish, so \(F(V)=0\).
No point on \(F(V)=\Lambda_h^4\) with \(\Lambda_h\neq0\) can therefore be
singular.  A smooth hypersurface in six complex affine coordinates has
complex dimension \(5\).

For the mass deformation, a general holomorphic quadratic mass source in the
four doublets is an antisymmetric tensor \(m_{IJ}V^{IJ}/2\).  If \(m\) is
nondegenerate, elementary alternating-form linear algebra gives a basis
\((e_1,e_2,e_3,e_4)\) with
\[
  m=m_1\,e^1\wedge e^2+m_2\,e^3\wedge e^4,
  \qquad m_1m_2\neq0;
\]
one chooses \(e_1\), then \(e_2\) with \(m(e_1,e_2)\neq0\), takes the
orthogonal complement of their span, and repeats in the remaining
two-dimensional nondegenerate alternating space.  In that Darboux basis, the
\(F\)-terms from
\eqref{eq:su2-nf2-constrained-mass-superpotential} are
\[
\begin{array}{lll}
  m_1+XV^{34}=0, & -XV^{24}=0, & XV^{23}=0,\\
  -XV^{13}=0, & XV^{14}=0, & m_2+XV^{12}=0,
\end{array}
\]
together with \(F(V)=\Lambda_h^4\).  If \(X=0\), the first and last equations
would force \(m_1=m_2=0\), contrary to the hypothesis.  Hence \(X\neq0\),
and the middle four equations set
\[
  V^{13}=V^{14}=V^{23}=V^{24}=0.
\]
The remaining two linear equations give
\[
  V^{12}=-\frac{m_2}{X},\qquad
  V^{34}=-\frac{m_1}{X}.
\]
The constraint then becomes
\[
  \frac{m_1m_2}{X^2}=\Lambda_h^4,
\]
so \(X=X_\pm\) and the two displayed solutions follow.  Substitution into
\eqref{eq:su2-nf2-diagonal-mass-source} gives
\[
  W_\pm
  =
  -\frac{2m_1m_2}{X_\pm}
  =
  \mp\,2\Lambda_h^2\sqrt{m_1m_2},
\]
as claimed.

Finally, the comparison with pure \(SU(2)\) is the standard holomorphic
threshold matching applied to two massive doublet pairs.  The one-loop
holomorphic exponents are \(b_0=4\) for \(SU(2)\) with two flavors and
\(b_0=6\) for pure \(SU(2)\).  In the Darboux mass convention the Pfaffian
of the antisymmetric mass matrix is \(m_1m_2\), so decoupling gives
\[
  \Lambda_0^6=m_1m_2\Lambda_h^4.
\]
Thus the two values above may be written as \(W_\pm=\mp2\Lambda_0^3\) after
choosing a cube-root branch for \(\Lambda_0^3\), exactly the two-branch
pure \(SU(2)\) chiral-source pattern.  This matching paragraph uses the
Wilsonian decoupling input; the smoothness and \(F\)-term solution were
already proved directly from the deformed Pfaffian equation.

\section{N Two Coulomb, Higgs, And Mixed Branches}

For \(\mathcal N=2\) theories, Coulomb-branch coordinates come from vector
multiplet scalars and have special K\"ahler geometry.  Higgs branches come from
hypermultiplet scalars and are hyperk\"ahler quotients:
\[
  \mathcal M_H=\mu_\mathbb R^{-1}(0)\cap\mu_\mathbb C^{-1}(0)/G .
\]
Mixed branches have both vector and hypermultiplet massless fields after
solving compatibility conditions between vector scalar expectation values
and hypermultiplet expectation values.  The Coulomb-branch metric may receive
quantum corrections encoded by Seiberg-Witten periods.  The phrase protected
Higgs-branch metric has a narrower meaning.

\paragraph{Smooth Wilsonian Higgs-branch protection.}
Let a Lagrangian theory with eight supercharges have compact gauge group
\(G\), hypermultiplet scalar space \(\mathcal X\), a supersymmetric Wilsonian
regulator, and declared mass and FI parameters.  Let
\(U\subset\mathcal M_H^{\rm sm}\) be a smooth open stratum of the Higgs branch
on which the gauge group is completely broken, vector multiplets and transverse
hypermultiplets have a mass gap \(M_\perp\), and no mixed branch or additional
massless sector meets \(U\).  The local two-derivative question is whether the
Wilsonian action for momenta \(|p|\ll M_\perp\), restricted to the massless
hypermultiplets on \(U\), contains a metric counterterm
\[
  \Delta S^{(2)}_H
  =
  \frac12\int d^dx\,
  \Delta g_{mn}(q)\,
  \partial_\mu q^m\partial^\mu q^n+\cdots .
\]
Eight supercharges require the completed torsion-free two-derivative sigma
model in \(4d\ \mathcal N=2\), \(3d\ \mathcal N=4\), or a declared
torsion-free \(2d\ \mathcal N=(4,4)\) reduction to have a hyperk\"ahler metric,
with the fermion and auxiliary couplings fixed by the Levi-Civita connection and
curvature.  In a general two-dimensional \((4,4)\) sigma model with
\(H=dB\), however, supersymmetry constrains \(g\) and \(H\) together through
torsionful connections; the target is HKT or bi-Hermitian rather than
hyperk\"ahler for the Levi-Civita connection.  Thus a genuine metric counterterm
is a local deformation of the intrinsic torsion-free metric on \(U\), not a
change of source coordinate or wave-function representative.

\begin{hypothesis}[Smooth torsion-free Higgs-branch metric-protection input]
\label{hyp:smooth-higgs-branch-metric-protection}
In a Wilsonian scheme preserving the eight supercharges, the relevant
\(R\)-symmetry, locality on \(U\), the transverse mass gap \(M_\perp\), and in
two dimensions a declared torsion-free \((4,4)\) hypermultiplet sigma-model
class, a local pure-Higgs two-derivative counterterm on a smooth fully Higgsed
stratum is cohomologous to a hypermultiplet coordinate change or to transport of
the declared FI/mass parameters.  Equivalently, after those source coordinates
have been matched, massive vector-sector and transverse-hypermultiplet loops do
not deform the intrinsic hyperk\"ahler metric on \(U\).
\end{hypothesis}

This is the theorem-boundary input used in this chapter.  Component
multiplicities, quotient dimension counts, and isolated heat-kernel traces are
only consistency checks for the input.  The Ward-identity content is local and
uses the multiplet type of every background.  The massless coordinates on \(U\)
sit in hypermultiplets, whereas gauge couplings, vector masses, and Coulomb
coordinates sit in vector-multiplet backgrounds, and FI parameters sit in their
corresponding triplet backgrounds.  A two-derivative pure Higgs metric term
must be an invariant made only from the massless hypermultiplet sigma-model
fields.  Once a vector or coupling spurion is inserted, the \(R\)-symmetry
quantum numbers and eight-supercharge measure move the operator into a
vector-sector term, a transport of the quotient level, or a higher-derivative
correction.
If a two-dimensional \((4,4)\) branch allows torsion \(H=dB\), the theorem
boundary is different: the local counterterm problem is the coupled deformation
of \((g,B,\Phi)\), or of the equivalent source multiplets, preserving the
torsionful \((4,4)\) connections.  Its metric projection is not governed by
Hypothesis~\ref{hyp:smooth-higgs-branch-metric-protection}.

The dimension-dependent reading of
Hypothesis~\ref{hyp:smooth-higgs-branch-metric-protection} is:
\begin{center}
\begin{tabular}{@{}p{0.22\linewidth}p{0.64\linewidth}@{}}
\textbf{dimension and algebra} & \textbf{local Wilsonian status}\\ \hline
\(4d\ \mathcal N=2\)
& Perturbatively, in a regulator preserving the eight supercharges and
\(SU(2)_R\), no local two-derivative counterterm depending on vector-multiplet
or gauge-coupling spurions deforms the intrinsic Higgs metric on \(U\).\\[0.4em]
\(3d\ \mathcal N=4\)
& The same local statement holds after dimensional reduction, with
\(SU(2)_H\times SU(2)_C\) separating hypermultiplet and vector-multiplet
backgrounds.  Monopole effects belong to the Coulomb-branch sector unless new
massless fields meet \(U\).\\[0.4em]
\(2d\ \mathcal N=(4,4)\)
& In a declared torsion-free class, the same local metric statement is the
dimensional reduction of the higher-dimensional hypermultiplet theorem.  If
\(H=dB\) is allowed, the supersymmetric target data are coupled:
\(g\), \(B\), and possible dilaton or source terms must satisfy the HKT or
bi-Hermitian \((4,4)\) constraints together.  Symmetrizing a trace-log kernel is
then only a raw tensor decomposition, not an independent metric theorem.
\end{tabular}
\end{center}
The global equality
\begin{equation}
  g_H
  =
  (g_{\mathcal X})_{{\rm HKQ}_{\zeta,G}}
  \label{eq:higgs-branch-protected-quotient-metric}
\end{equation}
is therefore a local theorem-boundary statement on a smooth fully Higgsed
stratum, after source matching, plus a separate nonperturbative continuum
assertion when the same stratum is known to avoid extra light sectors,
boundary degrees of freedom, and singular-throat contributions.

Here is the counterterm filter behind the local statement.  A Wilsonian
two-derivative correction on \(U\) can only fall into one of the following
classes:
\begin{center}
\begin{tabular}{@{}>{\raggedright\arraybackslash}p{0.24\linewidth}
>{\raggedright\arraybackslash}p{0.44\linewidth}
>{\raggedright\arraybackslash}p{0.22\linewidth}@{}}
\textbf{candidate} & \textbf{effect} & \textbf{intrinsic metric?}\\ \hline
hypermultiplet coordinate change
& changes \(g_{i\bar j}\) in a chosen chart or source normalization
& no\\[0.3em]
FI or mass-parameter transport
& moves the hyperk\"ahler quotient level or the chosen background
& not after matching\\[0.3em]
vector/coupling-spurion \(D\)-term
& carries vector-multiplet or Coulomb data; by the eight-supercharge Ward
identity it is a vector-sector or higher-derivative operator, not a pure Higgs
metric deformation
& no\\[0.3em]
torsionful two-dimensional \((4,4)\) deformation
& coupled \((\delta g,\delta B,\delta\Phi)\) or source deformation constrained
by torsionful target geometry, not an independent antisymmetric add-on
& outside the torsion-free metric theorem\\[0.3em]
singular or mixed-branch operator
& requires light fields not present in the sigma model on \(U\)
& outside
\end{tabular}
\end{center}
The filter is necessary rather than sufficient.  It says which local
functionals would count as intrinsic metric deformations; it does not evaluate
the determinant that would be needed for an independent background-field proof.

\begin{construction}[Ward/local-counterterm proof interface]
\label{constr:higgs-branch-ward-counterterm-interface}
The all-order statement must be a local cohomology statement, not a component
ledger.  Let \({\mathfrak F}^{(2)}_{H,U}\) be the space of local
two-derivative Wilsonian functionals on the smooth stratum \(U\), completed by
the fermion and auxiliary terms required by the eight supercharges and
depending only on the massless hypermultiplet background after the FI and mass
sources have been declared.  In two dimensions this space is the torsion-free
subspace unless the \((4,4)\) multiplet and regulator explicitly include
torsion.  Quotient this space by hypermultiplet field redefinitions and by
transport of the declared FI/mass coordinates.  The theorem one has to prove,
or quote with its hypotheses, is the vanishing
\begin{equation}
  H^0_{Q,{\rm loc}}\!\left({\mathfrak F}^{(2)}_{H,U}\right)_{\rm intrinsic}
  =
  0 .
  \label{eq:higgs-branch-local-counterterm-cohomology}
\end{equation}
Here \(Q\) denotes the preserved supercharge algebra together with the relevant
\(R\)-symmetry Ward identities, and ``intrinsic'' means after source matching
and coordinate changes have been removed.  Equation
\eqref{eq:higgs-branch-local-counterterm-cohomology} is the precise local
content of Hypothesis~\ref{hyp:smooth-higgs-branch-metric-protection}.

This interface has four checks.  First, the proof has to keep the source
multiplets typed: hypermultiplet coordinates, vector-multiplet or gauge
coupling spurions, FI triplets, and mass backgrounds are not interchangeable.
Second, it is local only on a fully Higgsed smooth stratum with a transverse
gap; a singular or mixed locus changes the low-energy field content.  Third, in
two dimensions the torsion-free class must be declared.  If \(H=dB\) is allowed,
the target of the proof is a coupled local deformation complex for
\((g,B,\Phi)\) preserving the torsionful \((4,4)\) connections; projecting a
generic kernel by index symmetry is not a supersymmetric classification.
Fourth, the regulator used in any determinant realization must preserve the same
\(Q\) and \(R\)-symmetry data used in the cohomology statement.

The rank-one trace-log construction below is therefore evidence for the local
theorem in one model: it exhibits how the gauge-fixed heavy rows, contacts,
zero-mode projection, and regulator can realize the cancellation.  It is not a
substitute for \eqref{eq:higgs-branch-local-counterterm-cohomology}.  In a
four-supercharge theory a local K\"ahler \(D\)-term can be a genuine metric
renormalization; if such a term is imported into the eight-supercharge Higgs
branch argument, the proof has silently changed the theory.
\end{construction}

\begin{construction}[Rank-one background-field derivation target]
\label{constr:higgs-branch-background-field-derivation-target}
For an explicit perturbative calculation, fix a four-dimensional
\(\mathcal N=2\) \(U(1)\) theory with two hypermultiplets, positive FI
parameter \(\zeta\), and a smooth patch \(q_1\ne0\) of
\(\{\mu_{\mathbb C}=0,\mu_{\mathbb R}=\zeta\}/U(1)\simeq T^\ast\mathbb P^1\).
Choose a supersymmetric background-field \(R_\xi\) gauge about a slowly varying
Higgs tangent background \(q_B(x)\).  The massive long multiplet has
\(M^2=g^2\zeta\) at constant \(q_B\), but the two-derivative coefficient is
the trace-log expansion
\begin{equation}
  \Pi_{mn}^{\rm 1-loop}
  =
  \left.
  \frac12\,{\rm STr}\!
  \left(
    {\mathcal O}^{-1}{\mathcal V}_{mn}
    -
    {\mathcal O}^{-1}{\mathcal V}_{m}
    {\mathcal O}^{-1}{\mathcal V}_{n}
  \right)
  \right|_{\partial q_B\,\partial q_B,\,{\rm local}},
  \label{eq:higgs-branch-background-field-trace-log}
\end{equation}
where \({\mathcal V}_{m}\) are the vertices linear in
\(\partial_\mu q_B^m\), and \({\mathcal V}_{mn}\) are seagull vertices
quadratic in the slow background.  At the level needed to define the
calculation, the component quadratic operator classes are
\[
\begin{aligned}
  {\mathcal O}_{a}^{\mu\nu}
  &=
  (-D_B^2+M^2)\delta^{\mu\nu}
  +(1-\xi^{-1})D_B^\mu D_B^\nu
  +{\mathcal R}_{a}^{\mu\nu}[q_B,\partial q_B],\\
  {\mathcal O}_{G}
  &=
  -D_B^2+\xi M^2+{\mathcal R}_{G}[q_B,\partial q_B],\\
  {\mathcal O}_{c}
  &=
  -D_B^2+\xi M^2+{\mathcal R}_{c}[q_B,\partial q_B],\\
  {\mathcal O}_{\phi,\rho}
  &=
  -D_B^2+M^2+{\mathcal R}_{\phi,\rho}[q_B,\partial q_B],\\
  {\mathcal D}_{F}^{\,2}
  &=
  -D_B^2+M^2+{\mathcal R}_{F}[q_B,\partial q_B].
\end{aligned}
\]
Here \(a_\mu\) is the gauge fluctuation, \(G\) the Goldstone/eaten
hypermultiplet combination, \(c,\bar c\) the ghosts, \(\phi,\rho\) the vector
scalar and radial hypermultiplet partners, and \({\mathcal D}_F\) the coupled
fermion operator.  A complete calculation must keep the nonminimal
\(D_B^\mu D_B^\nu\) term, the \(a_\mu\)-Goldstone mixing cancelled by the gauge
condition, all seagulls in \({\mathcal V}_{mn}\), the ghost determinant, and
the \(\xi\)-dependent cancellation before the limit of constant \(q_B\) is
taken.  It must also exhibit the gauge map, the gauge-fixing adjoint, the
projection away from tangent zero modes, and the supersymmetric heavy complex,
or an equivalent Ward identity, that pairs the tangent-dependent bosonic and
fermionic fluctuation vertices.  The same exercise in three or two dimensions
has to be repeated after dimensional reduction, because gauge components,
reduced vector scalars, and infrared/torsion data are different objects.
\end{construction}

\begin{example}[\(R_\xi\) determinant split in the rank-one Higgs patch]
\label{ex:higgs-branch-rxi-determinant-split}
The first nontrivial checkpoint in the calculation is already fixed by the
constant-background quadratic operators, and it is stricter than the old
component count.  At a point of the smooth patch take, in quotient coordinates,
\(q_B=(\sqrt\zeta,0)\) and choose the tangent \(u(x)\) transverse to the
infinitesimal gauge orbit.  Before expanding in \(\partial u\), the
gauge-fixed massive vector operator has the Euclidean momentum-space split
\begin{equation}
  {\mathcal O}_{a}^{\mu\nu}(p)
  =
  (p^2+M^2)P_T^{\mu\nu}
  +
  \xi^{-1}(p^2+\xi M^2)P_L^{\mu\nu},
  \qquad
  M^2=g^2\zeta ,
  \label{eq:higgs-branch-rxi-gauge-vector-split}
\end{equation}
where \(P_T\) and \(P_L\) are the transverse and longitudinal projectors.  The
field-independent factor \(\xi^{-1}\) in the longitudinal eigenvalue is not a
Higgs metric counterterm, so the local trace-log keeps the field-dependent
longitudinal factor \(p^2+\xi M^2\).  The Goldstone mode selected by the
gauge-fixing functional and the Faddeev--Popov ghost have the same
field-dependent operator,
\[
  {\mathcal O}_G(p)=p^2+\xi M^2,\qquad
  {\mathcal O}_c(p)=p^2+\xi M^2 .
\]
Hence the entire gauge-parameter-dependent determinant factor generated by
these operators is
\begin{equation}
  \frac12\log(p^2+\xi M^2)
  +
  \frac12\log(p^2+\xi M^2)
  -
  \log(p^2+\xi M^2)
  =
  0 .
  \label{eq:higgs-branch-rxi-xi-sector-cancellation}
\end{equation}
The three terms are, respectively, the longitudinal vector, the real Goldstone,
and the complex ghost pair.  Equation
\eqref{eq:higgs-branch-rxi-xi-sector-cancellation} evaluates the
\(\xi\)-dependent gauge-fixing factor.  The value of
\(\Pi_{mn}^{\rm 1-loop}\) requires evaluating and pairing the
tangent-dependent connection vertices, seagulls, fermions, vector scalar, radial
hypermultiplet fluctuation, and auxiliary contacts appearing in
\eqref{eq:higgs-branch-background-field-trace-log}.  What the split proves is
narrower but essential: any calculation that omits the ghost, the Goldstone, or
the nonminimal longitudinal vector mode has already lost gauge-parameter
independence.

The same split also shows why the dimensionally reduced cases cannot reuse the
four-dimensional \(A_\mu\) entry.  In \(d=4,3,2\) the vector contributes
\(d-1\) transverse factors and one longitudinal factor to
\eqref{eq:higgs-branch-rxi-xi-sector-cancellation}; the remaining \(4-d\)
components of the original connection are reduced vector-multiplet scalars with
\(\xi\)-independent operator \(p^2+M^2\) at this checkpoint.  The
two-dimensional calculation must still be supplemented by its separate
torsion/\(B\)-field and infrared regulator analysis.
\end{example}

\begin{example}[Frame-connection seagulls in the Higgs determinant]
\label{ex:higgs-branch-frame-connection-seagull}
The next checkpoint concerns the part of
\eqref{eq:higgs-branch-background-field-trace-log} generated only by changing
the orthonormal frame of massive fluctuations as the Higgs tangent moves.  Let
\({\mathcal H}_{\rm heavy}\) be one regulated heavy block, with the grading kept
fixed.  Suppose that along a path \(q_B(s)\) the corresponding quadratic
operator is only conjugated,
\[
  {\mathcal O}(s)=U(s){\mathcal O}_0U(s)^{-1},\qquad U(0)=1,
\]
and put \(T=\dot U(0)\).  Then
\[
  \dot{\mathcal O}=[T,{\mathcal O}_0],
  \qquad
  \ddot{\mathcal O}=[T,[T,{\mathcal O}_0]] .
\]
The second-order trace-log contribution from this pure frame connection is
\begin{equation}
  \frac12\,{\rm STr}\!
  \left(
    {\mathcal O}_0^{-1}\ddot{\mathcal O}
    -
    {\mathcal O}_0^{-1}\dot{\mathcal O}\,
    {\mathcal O}_0^{-1}\dot{\mathcal O}
  \right)
  =
  \frac12\,\left.\frac{d^2}{ds^2}{\rm STr}\log{\mathcal O}(s)\right|_{s=0}
  =
  0 .
  \label{eq:higgs-branch-frame-connection-seagull-identity}
\end{equation}
Thus the seagull \(\ddot{\mathcal O}\) and the double insertion
\(\dot{\mathcal O}\dot{\mathcal O}\) are not optional bookkeeping choices; they
are the two Taylor coefficients of the same covariant fluctuation operator.  In
the rank-one Higgs patch, the fixed-norm flavor rotation of the heavy
Goldstone/radial/vector-scalar/fermion frame has exactly this form before the
genuine curvature of the mass matrix is included.  Dropping the seagull would
manufacture a spurious Higgs-metric term even though the regulated determinant
has merely been written in a moving basis.

Equation~\eqref{eq:higgs-branch-frame-connection-seagull-identity} closes the
pure-frame part of the trace-log.  The non-frame pieces of the remainders in
the operators above carry the mass-curvature, auxiliary-contact, and
supersymmetric Ward-pairing data.  An evaluation of
\(\Pi_{mn}^{\rm 1-loop}\) has to sum those terms.
\end{example}

\begin{example}[Mass-curvature vertices and Ward pairing]
\label{ex:higgs-branch-mass-curvature-ward-pair}
After the pure frame-connection part has been removed, the next local datum is
the generated vertex pair rather than a field count.  For a heavy bosonic block
and the square of its fermionic partner write
\[
\begin{aligned}
  {\mathcal O}_B(s)
  &=
  {\mathcal O}_0+sX_B+\frac{s^2}{2}Y_B+O(s^3),\\
  {\mathcal D}_F(s)^2
  &=
  {\mathcal O}_0+sX_F+\frac{s^2}{2}Y_F+O(s^3).
\end{aligned}
\]
Their contribution to the second variation of the one-loop determinant is
\begin{equation}
\begin{aligned}
  \delta^2\Gamma_{B,F}
  &=
  \frac12\,{\rm Tr}\!
  \left(
    {\mathcal O}_0^{-1}Y_B
    -
    {\mathcal O}_0^{-1}X_B{\mathcal O}_0^{-1}X_B
  \right)\\
  &\quad
  -
  \frac12\,{\rm Tr}\!
  \left(
    {\mathcal O}_0^{-1}Y_F
    -
    {\mathcal O}_0^{-1}X_F{\mathcal O}_0^{-1}X_F
  \right).
  \label{eq:higgs-branch-mass-curvature-ward-pair}
\end{aligned}
\end{equation}
The Ward-paired cancellation is the statement that the regulator and auxiliary
completion identify the generated vertices,
\[
  X_B=X_F,\qquad Y_B=Y_F ,
\]
after all contact terms have been included.  Then
\(\delta^2\Gamma_{B,F}=0\) for that paired heavy block.  This is the determinant-level
replacement for the old multiplicity line: equal bosonic and fermionic
dimensions would not cancel
\eqref{eq:higgs-branch-mass-curvature-ward-pair} if, for example, the scalar
seagull or auxiliary contact shifts \(Y_B\), or a Yukawa/connection term shifts
\(X_F\).  In the rank-one Higgs calculation the radial hypermultiplet
fluctuation, vector scalar, heavy fermions, and auxiliary contacts must be
organized into these matched \(X,Y\) pairs before the residual
\(\Pi_{mn}^{\rm 1-loop}\) can be declared zero.
\end{example}

\begin{example}[Supercharge factorization of the heavy determinant]
\label{ex:higgs-branch-supercharge-factorization}
The previous checkpoint names the \(X,Y\) vertices that must be paired.  The
mechanism that can produce the pairing is the off-shell supersymmetric heavy
complex.  Work on one regulated massive block after the zero modes tangent to
the branch have been separated.  In a supersymmetric background-field gauge,
with auxiliary fields kept until the quadratic operator is assembled and with a
regulator preserving the chosen supercharge, the paired nonzero operators have
the form
\begin{equation}
  {\mathcal O}_B(s)
  =
  {\mathcal Q}(s)^\dagger{\mathcal Q}(s),
  \qquad
  {\mathcal O}_F(s)
  =
  {\mathcal Q}(s){\mathcal Q}(s)^\dagger .
  \label{eq:higgs-branch-supercharge-factorized-operators}
\end{equation}
Here \(s\) is again a coordinate along the slow Higgs tangent, and the finite
notation stands for the regulated elliptic heavy complex on the nonzero
spectrum.  If
\[
  {\mathcal Q}(s)
  =
  {\mathcal Q}_0+s{\mathcal Q}_1+\frac{s^2}{2}{\mathcal Q}_2+O(s^3),
\]
then expansion of \eqref{eq:higgs-branch-supercharge-factorized-operators}
generates
\begin{equation}
\begin{aligned}
  X_B
  &=
  {\mathcal Q}_1^\dagger{\mathcal Q}_0
  +
  {\mathcal Q}_0^\dagger{\mathcal Q}_1,
  &
  Y_B
  &=
  {\mathcal Q}_2^\dagger{\mathcal Q}_0
  +
  2{\mathcal Q}_1^\dagger{\mathcal Q}_1
  +
  {\mathcal Q}_0^\dagger{\mathcal Q}_2,\\
  X_F
  &=
  {\mathcal Q}_1{\mathcal Q}_0^\dagger
  +
  {\mathcal Q}_0{\mathcal Q}_1^\dagger,
  &
  Y_F
  &=
  {\mathcal Q}_2{\mathcal Q}_0^\dagger
  +
  2{\mathcal Q}_1{\mathcal Q}_1^\dagger
  +
  {\mathcal Q}_0{\mathcal Q}_2^\dagger .
  \label{eq:higgs-branch-supercharge-factorized-vertices}
\end{aligned}
\end{equation}
The middle terms \(2{\mathcal Q}_1^\dagger{\mathcal Q}_1\) and
\(2{\mathcal Q}_1{\mathcal Q}_1^\dagger\) are contact terms forced by squaring
the same first-order operator.  They are the finite-block model for the scalar
seagulls and auxiliary/Yukawa contacts that disappear if one writes only the
linear tangent vertices.

On the paired nonzero spectrum the determinants agree:
\begin{equation}
  \frac12\,\frac{d^2}{ds^2}{\rm Tr}\log
  ({\mathcal Q}^\dagger{\mathcal Q})
  -
  \frac12\,\frac{d^2}{ds^2}{\rm Tr}\log
  ({\mathcal Q}{\mathcal Q}^\dagger)
  =
  0 .
  \label{eq:higgs-branch-supercharge-factorized-supertrace}
\end{equation}
Thus the matched-vertex assertion in
\eqref{eq:higgs-branch-mass-curvature-ward-pair} is a spectral pairing
statement for the regulated heavy \(Q\)-complex, rather than a component count.
The rank-one Higgs calculation still has to obtain \({\mathcal Q}(s)\) from the
gauge-fixed Lagrangian and show that the chosen regulator acts on this paired
complex.  A regulator or auxiliary elimination that changes the bosonic
\(Y_B\) contact without the corresponding fermionic \(Y_F\) contact would change
\(\Pi_{mn}^{\rm 1-loop}\).
\end{example}

\begin{example}[Off-shell row completion of the Higgs heavy complex]
\label{ex:higgs-branch-off-shell-row-completion}
The factorized identity is useful for the Higgs-metric calculation only after
\({\mathcal Q}(s)\) has been assembled from the quadratic gauge-fixed
Lagrangian.  In the rank-one patch, after the branch tangent has been separated
from the massive fluctuations in an orthonormal frame adapted to \(q_B(s)\), the
regulated first-order map is built row by row:
\begin{equation}
  {\mathcal Q}(s)
  =
  \begin{pmatrix}
    {\mathcal Q}_{\rm kin}(s)\\
    {\mathcal Q}_{\rm gf}(s)\\
    {\mathcal Q}_{\rm mom}(s)\\
    {\mathcal Q}_{\rm Yuk}(s)
  \end{pmatrix}.
  \label{eq:higgs-branch-row-completed-heavy-map}
\end{equation}
The kinetic rows come from the covariant scalar, vector, and fermion kinetic
terms in the moving heavy frame.  The gauge-fixing rows come from the
background-field \(R_\xi\) functional
\[
  {\mathcal F}
  =
  D_B^\mu a_\mu
  -
  \xi g\,{\rm Im}(\bar q_B\,\delta q_{\rm eat}),
\]
together with the corresponding ghost linearization.  The moment-map rows are
the \(D,F\)-auxiliary square completions for the triplet moment-map fluctuation.
The Yukawa rows are the heavy gaugino-hypermultiplet fermion couplings generated
by the same background \(q_B(s)\).  These rows may be reorganized by field
redefinitions, but they cannot be deleted without retaining their Schur-complement
contact terms.

Writing
\[
  {\mathcal Q}_r(s)
  =
  {\mathcal Q}_{r,0}
  +
  s{\mathcal Q}_{r,1}
  +
  \frac{s^2}{2}{\mathcal Q}_{r,2}
  +
  O(s^3)
\]
for the row indexed by \(r\), the bosonic operator and its generated tangent
vertices are
\begin{equation}
\begin{aligned}
  {\mathcal O}_B
  &=
  \sum_r{\mathcal Q}_r^\dagger{\mathcal Q}_r,\\
  X_B
  &=
  \sum_r
  \left(
    {\mathcal Q}_{r,1}^\dagger{\mathcal Q}_{r,0}
    +
    {\mathcal Q}_{r,0}^\dagger{\mathcal Q}_{r,1}
  \right),\\
  Y_B
  &=
  \sum_r
  \left(
    {\mathcal Q}_{r,2}^\dagger{\mathcal Q}_{r,0}
    +
    2{\mathcal Q}_{r,1}^\dagger{\mathcal Q}_{r,1}
    +
    {\mathcal Q}_{r,0}^\dagger{\mathcal Q}_{r,2}
  \right).
  \label{eq:higgs-branch-row-completed-bosonic-vertices}
\end{aligned}
\end{equation}
The fermionic square has the same information with row indices left visible,
\begin{equation}
  (Y_F)_{rs}
  =
  {\mathcal Q}_{r,2}{\mathcal Q}_{s,0}^\dagger
  +
  2{\mathcal Q}_{r,1}{\mathcal Q}_{s,1}^\dagger
  +
  {\mathcal Q}_{r,0}{\mathcal Q}_{s,2}^\dagger ,
  \label{eq:higgs-branch-row-completed-fermionic-seagull}
\end{equation}
and similarly for \(X_F\).  Thus each row has a local square-completion contact
\[
  C_A
  =
  2\sum_{r\in A}{\mathcal Q}_{r,1}^\dagger{\mathcal Q}_{r,1}
\]
for any group \(A\) of gauge-fixing, moment-map, or Yukawa rows.  If those rows
are integrated out before the determinant is written, their Schur complement
must reproduce \(C_A\).  If the contact is simply dropped while the fermionic
operator is kept fixed, the second variation changes by
\begin{equation}
  \Pi_{\rm dropped}-\Pi_{\rm full}
  =
  -\frac12\,{\rm Tr}({\mathcal O}_0^{-1}C_A).
  \label{eq:higgs-branch-row-contact-residual}
\end{equation}
This is the component-level obstruction to treating the old long-multiplet
balance as a proof.  The actual rank-one determinant has to show that the
kinetic, gauge-fixing/ghost, moment-map auxiliary, and Yukawa rows assemble into
one regulated complex before it invokes the cancellation in
\eqref{eq:higgs-branch-supercharge-factorized-supertrace}.
\end{example}

\begin{example}[Rank-one row Jacobians from the component Lagrangian]
\label{ex:higgs-branch-rank-one-row-jacobians}
The row completion above is not meant as an abstract replacement for the
Lagrangian calculation.  In the \(U(1)\) model fixed in
Construction~\ref{constr:higgs-branch-background-field-derivation-target} the
first rows can be read directly from the gauge-fixed component action.  On the
zero-cotangent slice of the \(q_1\ne0\) patch write a fixed-norm quotient path
\[
  q_B(s)=\bigl(r(s),s\bigr),\qquad
  r(s)=(\zeta-s^2)^{1/2},\qquad
  \widetilde q_B(s)=0,
  \qquad
  q_B^\dagger\dot q_B=0 .
\]
Split a hypermultiplet fluctuation into real and imaginary parts
\(\delta q=\delta x+\ii\delta y\).  The component terms
\[
  \frac{g^2}{2}\mu_R^2
  +g^2|\mu_C|^2
  +\frac1{2\xi}{\mathcal F}^2
  +\sqrt2 g\,\bar q_B\lambda\psi+\hbox{h.c.}
\]
generate, before any determinant is taken, the row functionals
\begin{equation}
\begin{aligned}
  J_R(s)\delta x
  &=\sqrt2\,g\,q_B(s)^\dagger\delta x,\\
  J_{\rm gf}(s)\delta y
  &=\sqrt\xi\,g\,q_B(s)^\dagger\delta y,\\
  J_C^{(1,2)}(s)\delta\widetilde q
  &=g\,q_B(s)^\dagger
    \bigl(\delta\widetilde x,\delta\widetilde y\bigr),\\
  J_{\rm Yuk}(s)
  &\sim \sqrt2\,g\,q_B(s)^\dagger .
  \label{eq:higgs-branch-rank-one-row-jacobians}
\end{aligned}
\end{equation}
Here \(J_R\) is the real moment-map auxiliary row, \(J_C^{(1,2)}\) are the two
real rows of the complex moment map, \(J_{\rm gf}\) is the Higgs part of the
background \(R_\xi\) gauge functional, and \(J_{\rm Yuk}\) is the same
background vector appearing in the gaugino-hyperino mass row.  The constants
and spinor contractions differ, but the Higgs-background dependence is the
same row \(q_B(s)^\dagger\).

For any one of these rows with coefficient \(\alpha_A\), set
\[
  P_H(s)=\frac{q_B(s)q_B(s)^\dagger}{\zeta},
  \qquad
  P_T(s)=1-P_H(s),
  \qquad
  m_A^2=\alpha_A^2\zeta .
\]
Then the fixed-coordinate fluctuation operator generated by the row is
\begin{equation}
  J_A(s)^\dagger J_A(s)=m_A^2P_H(s),
  \qquad
  J_A(s)J_A(s)^\dagger=m_A^2,
  \qquad
  P_T(s)J_A(s)^\dagger=0 .
  \label{eq:higgs-branch-row-projector-mass}
\end{equation}
Thus the row has one massive normal eigenvalue and one quotient tangent zero
mode.  The nonzero eigenvalue is independent of \(s\); the \(s\)-dependence of
\(J_A^\dagger J_A\) is the rotation of the heavy projector \(P_H(s)\), not a
curvature of the physical mass.  This is the zero-mode separation that the old
long-multiplet balance did not contain.

Differentiating the row also displays the seagull/contact term that must travel
with the linear tangent vertex:
\begin{equation}
  \frac{d^2}{ds^2}\bigl(J_A^\dagger J_A\bigr)
  =
  \alpha_A^2
  \bigl(
    \ddot q_B q_B^\dagger
    +2\dot q_B\dot q_B^\dagger
    +q_B\ddot q_B^\dagger
  \bigr),
  \qquad
  \frac{d^2}{ds^2}\bigl(J_AJ_A^\dagger\bigr)=0 .
  \label{eq:higgs-branch-row-jacobian-contact}
\end{equation}
The middle term \(2\dot q_B\dot q_B^\dagger\) is the row-level
square-completion contact already abstracted in
\eqref{eq:higgs-branch-row-completed-bosonic-vertices}.  Omitting it turns the
motion of the projector into a fake mass-curvature contribution.  Keeping it
shows that the moment-map, gauge-fixing, and Yukawa rows derived from the
component Lagrangian contribute only the heavy-frame determinant cell until a
genuine variation of a nonzero mass or an unpaired auxiliary/Yukawa contact is
introduced.
\end{example}

\begin{example}[Gauge-fixed heavy package and zero-mode projection]
\label{ex:higgs-branch-gauge-fixed-heavy-package}
The row Jacobians still do not define a determinant until they are assembled
with the gauge map and the tangent projection.  In the rank-one \(U(1)\) model,
the infinitesimal gauge action on a fluctuation about \(q_B,\widetilde q_B\) is
\begin{equation}
  {\mathcal R}_B\epsilon
  =
  \bigl(D_{B,\mu}\epsilon,\,
        \ii g\epsilon q_B,\,
        -\ii g\epsilon\widetilde q_B\bigr),
  \label{eq:higgs-branch-gauge-map}
\end{equation}
where the three entries are the connection, \(q\), and \(\widetilde q\)
directions.  The background \(R_\xi\) gauge-fixing row is the adjoint companion,
with the \(\xi\)-weight included,
\begin{equation}
  {\mathcal G}_{B,\xi}(a,\eta_q,\eta_{\widetilde q})
  =
  D_B^\mu a_\mu
  -
  \xi g\,{\rm Im}
  \bigl(q_B^\dagger\eta_q
        -
        \widetilde q_B\eta_{\widetilde q}^\dagger\bigr).
  \label{eq:higgs-branch-gauge-fixing-adjoint-row}
\end{equation}
Linearizing this row along the same gauge orbit gives the ghost operator
\begin{equation}
  {\mathcal O}_c
  =
  {\mathcal G}_{B,\xi}{\mathcal R}_B
  =
  -D_B^2
  +
  \xi g^2\bigl(|q_B|^2+|\widetilde q_B|^2\bigr)
  +
  {\mathcal V}_c[q_B,\partial q_B].
  \label{eq:higgs-branch-ghost-from-gauge-map}
\end{equation}
Thus the ghost row is not a separately chosen determinant.  Its mass and
tangent vertices are generated by the same \(q_B\)-dependent map that produces
the longitudinal-vector and Goldstone rows.  On the zero-cotangent fixed-norm
slice, \(|q_B|^2=\zeta\), this reproduces the common \(p^2+\xi M^2\) factor in
\eqref{eq:higgs-branch-rxi-xi-sector-cancellation}.

The bosonic heavy rows are then the gauge-fixed package
\begin{equation}
  {\mathcal Q}_{B,\xi}
  =
  \begin{pmatrix}
    {\mathcal G}_{B,\xi}\\
    d\mu_{\mathbb R,B}\\
    d\mu_{\mathbb C,B}\\
    {\mathcal K}_{B}\\
    {\mathcal Y}_{B}
  \end{pmatrix},
  \label{eq:higgs-branch-gauge-fixed-heavy-package}
\end{equation}
where \({\mathcal K}_B\) denotes the kinetic moving-frame rows and
\({\mathcal Y}_B\) the Yukawa mass rows.  The tangent modes of the Higgs branch
are the simultaneous zero modes
\begin{equation}
  {\mathcal T}_B U
  =
  \{\tau:\ d\mu_{\mathbb R,B}\tau=0,
          d\mu_{\mathbb C,B}\tau=0,
          {\mathcal G}_{B,\xi}(0,\tau)=0\}/{\rm im}\,{\mathcal R}_B .
  \label{eq:higgs-branch-tangent-zero-mode-space}
\end{equation}
Let \(P_{\rm nz}=1-P_{{\mathcal T}_B U}\) be the projection onto the nonzero
heavy complement.  The operator inverted in the local trace-log is
\begin{equation}
  {\mathcal O}_{B,{\rm nz}}
  =
  P_{\rm nz}\,
  {\mathcal Q}_{B,\xi}^\dagger{\mathcal Q}_{B,\xi}\,
  P_{\rm nz},
  \qquad
  {\mathcal O}_{F,{\rm nz}}
  =
  P_{\rm nz}^{F}\,
  {\mathcal Q}_{F,\xi}{\mathcal Q}_{F,\xi}^\dagger\,
  P_{\rm nz}^{F}.
  \label{eq:higgs-branch-projected-heavy-operators}
\end{equation}
Here \({\mathcal Q}_{F,\xi}\) is the fermionic row map fixed by the same
background gauge choice and preserved supercharge.
At nonzero momentum, and on a fully Higgsed smooth stratum, the principal symbol
of the gauge-fixed package is elliptic on this complement: the gauge row pairs
longitudinal gauge and Goldstone directions, the moment-map rows pair the normal
hypermultiplet directions, the Yukawa rows pair the corresponding fermions, and
the only excluded zero modes are the branch tangents.  Omitting \(P_{\rm nz}\)
makes \({\mathcal O}^{-1}\) in
\eqref{eq:higgs-branch-background-field-trace-log} ill-defined; choosing a ghost
operator not equal to \({\mathcal G}_{B,\xi}{\mathcal R}_B\) destroys the
\(\xi\)-cancellation; and deleting a moment-map or Yukawa row removes the contact
terms displayed in
\eqref{eq:higgs-branch-row-completed-bosonic-vertices}.
\end{example}

\begin{example}[Regulator-compatible heavy-complex supertrace]
\label{ex:higgs-branch-regulator-compatible-heavy-complex}
The gauge-fixed package above still has to be inserted into a regulator that
acts on the same heavy complex.  On the nonzero spectrum of
\({\mathcal Q}(s)\), choose a Wilsonian or Pauli--Villars regulator whose
mass parameters are tangent-independent supersymmetric spurions and whose
bosonic and fermionic regulator multiplets use the same function
\[
  F_\Lambda(\lambda)
  =
  \sum_{\alpha=0}^N \eta_\alpha
  \log(\lambda+M_\alpha^2),
  \qquad
  \eta_0=1,\quad M_0=0 .
\]
The usual short-distance subtraction conditions are imposed on the same list
of regulator multiplets, for example
\[
  \sum_\alpha\eta_\alpha=0,\qquad
  \sum_\alpha\eta_\alpha M_\alpha^2=0
\]
at the level relevant for the local two-derivative coefficient.  The regulated
heavy determinant cell is then
\begin{equation}
  \Gamma_{\Lambda,{\rm heavy}}(s)
  =
  \frac12{\rm Tr}_{B,{\rm nz}}
    F_\Lambda({\mathcal Q}(s)^\dagger{\mathcal Q}(s))
  -
  \frac12{\rm Tr}_{F,{\rm nz}}
    F_\Lambda({\mathcal Q}(s){\mathcal Q}(s)^\dagger).
  \label{eq:higgs-branch-regulated-heavy-supertrace}
\end{equation}
Because the two nonzero spectra are identical, every paired eigenvalue
\(\lambda_a(s)>0\) gives
\[
  \frac12F_\Lambda(\lambda_a(s))
  -
  \frac12F_\Lambda(\lambda_a(s))=0
\]
before expanding in the slow tangent.  Hence
\begin{equation}
  \left.\frac{d^2}{ds^2}
  \Gamma_{\Lambda,{\rm heavy}}(s)\right|_{s=0}
  =
  0
  \label{eq:higgs-branch-regulated-heavy-supertrace-variation}
\end{equation}
for the paired heavy block.  This is the regulator version of
\eqref{eq:higgs-branch-supercharge-factorized-supertrace}: the cancellation is
not a cancellation of bare component multiplicities, but a cancellation of the
same regulated spectral function on the same nonzero supersymmetric complex.

Three failures are now visible.  First, if tangent zero modes are not projected
out before \(F_\Lambda\) is applied, the logarithm is not a local heavy-field
determinant.  Second, if the bosonic and fermionic regulator masses differ,
the second variation contains
\[
  \frac12\sum_a
  \left.\frac{d^2}{ds^2}
  \bigl(F_{\Lambda,B}(\lambda_a(s))
        -F_{\Lambda,F}(\lambda_a(s))\bigr)\right|_{s=0},
\]
which is a regulator artifact in the Higgs metric.  Third, if a
Pauli--Villars mass is allowed to depend on the Higgs coordinate without being
part of the same supersymmetric source multiplet, the regulator itself has
inserted a spurious background vertex.  A complete rank-one calculation must
therefore specify the regulator as part of the heavy \(Q\)-complex and only
then take the local trace-log coefficient in
\eqref{eq:higgs-branch-background-field-trace-log}.
\end{example}

\begin{example}[Metric extraction after source transport]
\label{ex:higgs-branch-metric-transport-projection}
The regulated heavy determinant in
\eqref{eq:higgs-branch-regulated-heavy-supertrace} is still one step away from
an intrinsic Higgs-metric statement.  Its two-derivative output is a local
symmetric kernel
\begin{equation}
  \Delta\Gamma_{\Lambda,H}^{(2)}
  =
  \frac12\int d^dx\,
  \Pi_{\Lambda,mn}(q)\,
  \partial_\mu q^m\partial^\mu q^n ,
  \label{eq:higgs-branch-metric-kernel-output}
\end{equation}
and the intrinsic part is obtained only after quotienting by hypermultiplet
field redefinitions and by the declared FI/mass source transport:
\begin{equation}
  [\Pi_\Lambda]_{\rm int}
  \in
  \frac{J^2{\rm Sym}^2T^\ast U}
       {J^2\{\mathcal L_v g_H\}
        +
        {\rm span}\{J^2(\partial_{\lambda^A}g_H)\}} .
  \label{eq:higgs-branch-metric-kernel-quotient}
\end{equation}
Here \(J^2\) denotes the local two-jet needed to test the metric and curvature
channel at the chosen point of the smooth stratum.  Thus a nonzero
\(\Pi_{\Lambda,mn}\) is not automatically a new Higgs metric, and a zero
component supertrace is not by itself enough: the determinant must land in the
redundant subspace displayed in
\eqref{eq:higgs-branch-metric-kernel-quotient}.

In the rank-one quotient this can be checked on the zero-cotangent
\(\mathbb P^1\) slice.  After fixing the normal-coordinate representative at
\(z=0\), write \(r=|z|^2\).  The quotient metric restricts to
\[
  g_\zeta
  =
  \zeta(1+r)^{-2}\,dz\,d\bar z
  =
  \zeta(1-2r+3r^2+O(r^3))\,dz\,d\bar z .
\]
The FI-source direction in this two-jet is therefore
\begin{equation}
  J^2(\partial_\zeta g_\zeta)=(1,-2,3).
  \label{eq:higgs-branch-cp1-slice-source-jet}
\end{equation}
If the trace-log kernel produces
\[
  J^2\Pi_\Lambda=a(1,-2,3),
\]
then it is only a transport of the quotient level \(\zeta\), not a new local
metric.  If instead
\begin{equation}
  J^2\Pi_\Lambda=a(1,-2,3)+(0,0,\varepsilon),
  \qquad \varepsilon\ne0,
  \label{eq:higgs-branch-intrinsic-jet-residual}
\end{equation}
the point metric and the long-multiplet inventory can still look harmless, but
the curvature two-jet contains an intrinsic residual.  This is the local
observable the Ward/cohomology theorem must kill.  The background-field
calculation is therefore complete only after it has projected its
\(\Pi_{\Lambda,mn}\) through
\eqref{eq:higgs-branch-metric-kernel-quotient}; otherwise one has checked a
determinant identity without extracting the branch sigma-model datum.
\end{example}

\begin{example}[Two-jet quotient by Higgs field redefinitions]
\label{ex:higgs-branch-field-redefinition-two-jet-quotient}
The quotient in \eqref{eq:higgs-branch-metric-kernel-quotient} removes local
hypermultiplet coordinate changes as well as source transport.  The rank-one
\(\mathbb P^1\) slice gives a finite version of this subtraction.  Keep
\(r=|z|^2\), and let a \(U(1)\)-invariant infinitesimal change of Higgs
coordinate be
\[
  v^z=(a+br)z,\qquad v^{\bar z}=(a+br)\bar z .
\]
For
\[
  g_\zeta=\zeta(1-2r+3r^2+O(r^3))\,dz\,d\bar z,
\]
the Lie derivative has two-jet
\begin{equation}
  J^2(\mathcal L_v g_\zeta)
  =
  \zeta\,\bigl(2a,\,-8a+4b,\,18a-12b\bigr).
  \label{eq:higgs-branch-field-redefinition-two-jet}
\end{equation}
Modulo the FI-source direction
\[
  S=J^2(\partial_\zeta g_\zeta)=(1,-2,3),
\]
the independent field-redefinition direction can be represented by
\[
  B_{\rm fr}=(0,4,-12).
\]
Thus a local trace-log output
\[
  K=(k_0,k_1,k_2)
\]
has, after source matching and Higgs-coordinate quotienting, the intrinsic
two-jet representative
\begin{equation}
  K_{\rm int}
  =
  \bigl(0,0,k_2+3k_1+3k_0\bigr).
  \label{eq:higgs-branch-two-jet-intrinsic-residual}
\end{equation}
Both \(S\) and \(B_{\rm fr}\) have zero residual in
\eqref{eq:higgs-branch-two-jet-intrinsic-residual}.  Therefore a determinant
kernel of the form
\[
  K=\alpha S+\beta B_{\rm fr}
\]
is a mixture of FI transport and a coordinate representative, while
\[
  K=\alpha S+\beta B_{\rm fr}+(0,0,\varepsilon)
\]
contains an intrinsic curvature-channel residual \(\varepsilon\).  A
source-only subtraction would wrongly flag the \(\beta B_{\rm fr}\) term as a
metric correction; a point-metric comparison would miss the
\((0,0,\varepsilon)\) term.  The physical output of the heavy determinant is
the residual class in
\eqref{eq:higgs-branch-two-jet-intrinsic-residual}, not the unprojected
two-jet coefficients.
\end{example}

\begin{example}[Dimensional reduction of the Higgs heavy rows]
\label{ex:higgs-branch-dimensional-reduction-row-audit}
The row completion also explains why the same four-dimensional component table
cannot be reused unchanged in lower dimensions.  Reduce the rank-one
calculation on \(4-d\) flat directions and restrict to backgrounds independent
of those directions.  The four-dimensional connection decomposes as
\[
  A_M=(A_\mu,\sigma_I),
  \qquad
  \mu=1,\ldots,d,\quad I=1,\ldots,4-d .
\]
At a constant Higgs background the gauge-fixed \(d\)-dimensional connection
keeps the split
\[
  {\mathcal O}_{a,d}^{\mu\nu}(p)
  =
  (p_d^2+M^2)P_{T,d}^{\mu\nu}
  +
  \xi^{-1}(p_d^2+\xi M^2)P_{L,d}^{\mu\nu},
\]
while each reduced connection scalar has
\[
  {\mathcal O}_{\sigma_I}(p)=p_d^2+M^2 .
\]
The \(\xi\)-dependent determinant still cancels dimension by dimension,
\[
  \frac12\log(p_d^2+\xi M^2)
  +
  \frac12\log(p_d^2+\xi M^2)
  -
  \log(p_d^2+\xi M^2)
  =
  0,
\]
but the remaining \(4-d\) fields are not longitudinal gauge modes.  They are
reduced vector-multiplet scalars and their tangent vertices must be paired in
the lower-dimensional heavy complex.

In row language this means that the reduced theory has extra rows
\[
  {\mathcal Q}_{\sigma_I}(s)
  =
  {\mathcal Q}_{\sigma_I,0}
  +
  s{\mathcal Q}_{\sigma_I,1}
  +
  \frac{s^2}{2}{\mathcal Q}_{\sigma_I,2}
  +
  O(s^3),
  \qquad I=1,\ldots,4-d .
\]
Their square-completion contribution is
\begin{equation}
  C_{\rm red}^{(d)}
  =
  2\sum_{I=1}^{4-d}
  {\mathcal Q}_{\sigma_I,1}^\dagger{\mathcal Q}_{\sigma_I,1}.
  \label{eq:higgs-branch-reduced-scalar-row-contact}
\end{equation}
If the reduced scalars are hidden inside a four-dimensional gauge-field entry
and their contact is dropped, the local second variation shifts by
\begin{equation}
  \Pi_{\rm omit\,red}^{(d)}-\Pi_{\rm full}^{(d)}
  =
  -\frac12{\rm Tr}
  \left({\mathcal O}_0^{-1}C_{\rm red}^{(d)}\right).
  \label{eq:higgs-branch-reduced-scalar-contact-residual}
\end{equation}
Thus the dimensionally reduced diagnostic count is at most
\begin{equation}
  {\mathcal B}_{\rm long}^{d}
  =
  d+(4-d)+2+4-2-8
  =
  0,
  \label{eq:higgs-branch-d-dimensional-balance-diagnostic}
\end{equation}
where the first term is the \(d\)-dimensional connection and the second is the
reduced scalar package.  The equality is not the determinant calculation.  It is
a check that the field inventory has not changed after the row contacts in
\eqref{eq:higgs-branch-reduced-scalar-row-contact} have been included.

In \(3d\ \mathcal N=4\), the \(\sigma_I\) fields belong to the vector-multiplet
sector separated by \(SU(2)_H\times SU(2)_C\).  In a torsion-free
\(2d\ \mathcal N=(4,4)\) reduction, the raw two-index kernel may be decomposed
as
\[
  \Pi_{mn}^{(g)}
  =
  \frac12(\Pi_{mn}+\Pi_{nm}),
  \qquad
  \Pi_{mn}^{(B)}
  =
  \frac12(\Pi_{mn}-\Pi_{nm}),
\]
but only \(\Pi^{(g)}\) belongs to the torsion-free metric theorem.  If the
two-dimensional branch allows \(H=dB\), the supersymmetric counterterm is instead
a coupled package
\[
  \delta{\mathcal C}_{(4,4)}
  =
  (\delta g_{mn},\delta B_{mn},\delta\Phi;\delta\nabla^{(+)},
  \delta\nabla^{(-)}),
  \qquad
  \nabla^{(\pm)}
  =
  \nabla^g\pm\frac12 g^{-1}H .
\]
A nonzero antisymmetric component is then not something to discard after
symmetrization; it signals that the HKT or bi-Hermitian deformation complex has
to be solved.  The metric-protection claim never follows from recycling the
four-dimensional \(A_\mu\) row.
\end{example}

After the local cohomology statement
\eqref{eq:higgs-branch-local-counterterm-cohomology}, the corresponding Ward
identity, or an explicit evaluation of
\eqref{eq:higgs-branch-background-field-trace-log} has paired these operators,
one may record the residual long-multiplet balance as a quick diagnostic.  In a
four-dimensional constant-mass patch the signed component balance is
\begin{center}
\begin{tabular}{@{}lrrrrrrr@{}}
\textbf{component}
& \(A_\mu\) & \(\phi_V\) & \(q_{\rm eaten}\)
& \(c,\bar c\) & \(D\) & \(\psi\) & \textbf{sum}\\ \hline
\textbf{coefficient}
& \(4\) & \(2\) & \(4\) & \(-2\) & \(0\) & \(-8\) & \(0\)
\end{tabular}
\end{center}
\begin{equation}
  {\mathcal B}_{\rm long}^{4d}
  =
  4+2+4-2+0-8
  =
  0 .
  \label{eq:higgs-branch-heavy-multiplet-balance-heuristic}
\end{equation}
This line computes only the signed long-multiplet balance.  It is useful for
detecting missing ghosts, missing eaten partners, or mass splitting, but the
vanishing of \(\Pi_{mn}^{\rm 1-loop}\) is a stronger statement about
\eqref{eq:higgs-branch-background-field-trace-log}.  The calculation is local
on \(U\); at a point where the long multiplet becomes massless, the branch EFT
has changed.

The theorem-status bookkeeping is therefore:
\begin{center}
\begin{tabular}{@{}lp{0.58\linewidth}@{}}
\textbf{smooth Higgs stratum} & local metric protection by
Hypothesis~\ref{hyp:smooth-higgs-branch-metric-protection}; the counterterm
filter and long-multiplet balance are consistency tests,\\[0.2em]
\textbf{coordinate representative} & scheme-dependent components of the same
intrinsic metric,\\[0.2em]
\textbf{global nonperturbative equality} & additional continuum assertion,
valid only after singular, boundary, and mixed-branch sectors are excluded,\\[0.2em]
\textbf{Coulomb branch} & quantum-corrected vector-multiplet metric in
general,\\[0.2em]
\textbf{singular or mixed locus} & new light fields; branch sigma-model metric
no longer closed by itself.
\end{tabular}
\end{center}

The quotient statement has concrete content already in the simplest
hypermultiplet model.  We spell it out because this is the local model behind
many later Higgs-branch calculations.

\begin{proposition}[Rank-one hypermultiplet quotient]
\label{prop:n2-rank-one-hyperkahler-quotient}
Let \(N\geq2\).  Let \(U(1)\) act on
\[
  V=\mathbb C^N_q\oplus\mathbb C^N_{\widetilde q}
\]
by
\[
  q_i\longmapsto \ee^{\ii\alpha}q_i,\qquad
  \widetilde q_i\longmapsto
  \ee^{-\ii\alpha}\widetilde q_i,\qquad i=1,\ldots,N .
\]
Use the complex structure in which \(q_i,\widetilde q_i\) are holomorphic
coordinates.  The complex moment map and real moment map are
\[
  \mu_{\mathbb C}(q,\widetilde q)
  =
  \sum_{i=1}^N q_i\widetilde q_i,
  \qquad
  \mu_{\mathbb R}(q,\widetilde q)
  =
  \sum_{i=1}^N |q_i|^2-\sum_{i=1}^N|\widetilde q_i|^2 .
\]
For real FI parameter \(\zeta>0\) and complex FI parameter \(0\), the
Higgs-branch quotient
\begin{equation}
  \mathcal M_H(\zeta)
  =
  \{\,\mu_{\mathbb C}=0,\ \mu_{\mathbb R}=\zeta\,\}/U(1)
  \label{eq:rank-one-hyperkahler-quotient}
\end{equation}
is biholomorphic, in this complex structure, to \(T^\ast\mathbb P^{N-1}\).
Its complex dimension is \(2(N-1)\), equivalently its real dimension is
\(4(N-1)\).  On the patch \(U_a=\{q_a\neq0\}\), local holomorphic
coordinates are
\begin{equation}
  z_i^{(a)}=\frac{q_i}{q_a},\qquad
  p_i^{(a)}=q_a\widetilde q_i,\qquad i\neq a,
  \label{eq:rank-one-hyperkahler-local-coordinates}
\end{equation}
and the complex moment-map equation determines
\begin{equation}
  q_a\widetilde q_a
  =
  -\sum_{i\neq a}z_i^{(a)}p_i^{(a)} .
  \label{eq:rank-one-hyperkahler-moment-solve}
\end{equation}
The holomorphic one-form
\[
  \Theta=\sum_{i=1}^N\widetilde q_i\,dq_i
\]
descends to the canonical cotangent one-form
\begin{equation}
  \Theta=\sum_{i\neq a}p_i^{(a)}\,dz_i^{(a)}
  \label{eq:rank-one-hyperkahler-cotangent-one-form}
\end{equation}
on \(U_a\).
\end{proposition}

\begin{proof}
First remove the real quotient from the holomorphic bookkeeping.  If
\(\mu_{\mathbb R}=\zeta>0\), then \(q\neq0\).  Conversely, fix a point
\((q,\widetilde q)\) with \(q\neq0\) and \(\mu_{\mathbb C}=0\).  Along the
positive real part of the complexified orbit,
\[
  q\mapsto rq,\qquad
  \widetilde q\mapsto r^{-1}\widetilde q,\qquad r>0,
\]
the real moment map becomes
\[
  f(r)=r^2\|q\|^2-r^{-2}\|\widetilde q\|^2 .
\]
This function is continuous and strictly increasing on \(r>0\), with limit
\(-\infty\) as \(r\to0\) if \(\widetilde q\neq0\) and limit \(0\) if
\(\widetilde q=0\), and with limit \(+\infty\) as \(r\to\infty\).  Hence
there is a unique positive \(r\) solving \(f(r)=\zeta\).  The remaining phase
is precisely the compact \(U(1)\) action.  Thus the quotient
\eqref{eq:rank-one-hyperkahler-quotient} is identified with the complex
quotient
\begin{equation}
  \{\,\mu_{\mathbb C}=0,\ q\neq0\,\}/\mathbb C^\ast .
  \label{eq:rank-one-hyperkahler-complex-quotient}
\end{equation}

On \(U_a\), the coordinates in
\eqref{eq:rank-one-hyperkahler-local-coordinates} are invariant under
\(\mathbb C^\ast\).  They determine the orbit because \(q_a\) supplies the
complexified gauge scale, the ratios determine all \(q_i\), the \(p_i^{(a)}\)
determine all \(\widetilde q_i\) with \(i\neq a\), and
\eqref{eq:rank-one-hyperkahler-moment-solve} determines
\(\widetilde q_a\).  There are \(2(N-1)\) complex coordinates, so the complex
dimension is \(2(N-1)\).

It remains to check that the local charts glue as the cotangent bundle, not
just as an abstract \(2(N-1)\)-dimensional quotient.  It suffices to compare
two patches; relabel the indices as \(a=1\) and \(b=2\).  Write
\[
  z_i=\frac{q_i}{q_1},\quad p_i=q_1\widetilde q_i
  \quad (i\neq1),
\]
and on the patch \(q_2\neq0\),
\[
  w_1=\frac{q_1}{q_2}=\frac1{z_2},\qquad
  w_i=\frac{q_i}{q_2}=\frac{z_i}{z_2}\quad (i\neq1,2),
\]
\[
  r_1=q_2\widetilde q_1=-z_2\sum_{i\neq1}z_ip_i,
  \qquad
  r_i=q_2\widetilde q_i=z_2p_i\quad (i\neq1,2),
\]
where the formula for \(r_1\) uses \(\mu_{\mathbb C}=0\).  A direct
differentiation gives
\begin{equation}
  r_1\,dw_1+\sum_{i\neq1,2}r_i\,dw_i
  =
  \sum_{i\neq1}p_i\,dz_i .
  \label{eq:rank-one-hyperkahler-one-form-transition}
\end{equation}
Thus the transition functions preserve the canonical holomorphic cotangent
one-form.

Finally, on a single patch,
\[
  dq_i=z_i^{(a)}\,dq_a+q_a\,dz_i^{(a)}\qquad (i\neq a).
\]
Using \eqref{eq:rank-one-hyperkahler-moment-solve}, the coefficient of
\(dq_a\) in \(\sum_i\widetilde q_i\,dq_i\) is
\[
  \widetilde q_a+\sum_{i\neq a}z_i^{(a)}\widetilde q_i=0,
\]
while the coefficient of \(dz_i^{(a)}\) is
\(q_a\widetilde q_i=p_i^{(a)}\).  This proves
\eqref{eq:rank-one-hyperkahler-cotangent-one-form} and identifies the quotient
with \(T^\ast\mathbb P^{N-1}\).
\end{proof}

\begin{example}[Two-dimensional \(\mathcal N=(4,4)\) ADHM Higgs branch]
\label{ex:d1-d5-adhm-higgs-branch}
The standard gauge-theory model behind the D1--D5 Higgs branch can be stated
without using any string-theory input.  Fix
\[
  K\simeq\mathbb C^{Q_1},\qquad W\simeq\mathbb C^{Q_5},
\]
and take a two-dimensional \(\mathcal N=(4,4)\) gauge theory with gauge group
\(U(K)\).  Its hypermultiplet scalars consist of an adjoint hypermultiplet
\((B_1,B_2)\in{\rm End}(K)^{\oplus2}\) and a fundamental hypermultiplet
\[
  I\in{\rm Hom}(W,K),\qquad J\in{\rm Hom}(K,W).
\]
In one complex structure the complex and real moment-map equations are
\begin{align}
  [B_1,B_2]+IJ&=0,
  \label{eq:d1-d5-adhm-complex-moment}\\
  [B_1,B_1^\dagger]+[B_2,B_2^\dagger]+II^\dagger-J^\dagger J
  &=\zeta\,{\bf 1}_K .
  \label{eq:d1-d5-adhm-real-moment}
\end{align}
For \(\zeta>0\), the Higgs branch is the hyperk\"ahler quotient
\begin{equation}
  \mathcal M_{Q_1,Q_5}(\zeta)
  =
  \{\,\eqref{eq:d1-d5-adhm-complex-moment},
       \eqref{eq:d1-d5-adhm-real-moment}\,\}/U(K),
  \label{eq:d1-d5-adhm-higgs-quotient}
\end{equation}
equivalently the stable complex quotient of
\([B_1,B_2]+IJ=0\) by \(GL(K)\).  The finite-dimensional count is already
physical:
\[
\begin{array}{rcl}
\dim_{\mathbb C}\{B_1,B_2,I,J\}&=&2Q_1^2+2Q_1Q_5,\\
\dim_{\mathbb C}\mu_{\mathbb C}^{-1}(0)&=&Q_1^2,\\
\dim_{\mathbb C}GL(K)&=&Q_1^2,
\end{array}
\qquad
\dim_{\mathbb R}\mathcal M_{Q_1,Q_5}=4Q_1Q_5 .
\]
On a smooth stable stratum where \(U(K)\) is completely broken and the local
Higgs-branch sigma model is taken in the torsion-free \((4,4)\) class, the
two-derivative metric is the quotient metric in the smooth Wilsonian protection
statement above.  This is the protected Higgs-branch metric statement.  It is
stronger than saying that the algebraic quotient has the ADHM equations,
because it identifies the Wilsonian kinetic term of the massless QFT fields.
It is also weaker than a
claim about every point of the quotient: when \(\zeta=0\), the locus
\(I=J=0\) and commuting \(B_1,B_2\) meets the small-instanton and Coulomb
directions, the gauge stabilizer can jump, and additional massless fields must
be restored.  The Coulomb branch of the same two-dimensional gauge theory is a
boundary case rather than another protected metric; integrating out
hypermultiplets changes the vector-multiplet metric, so the quotient
calculation alone does not determine the full infrared geometry.
Nor does the quotient metric by itself construct the noncompact throat, the
boundary-condition data, or the interacting infrared SCFT.  Those require a
separate continuum argument.  The protected statement here is only the
torsion-free local two-derivative metric on the smooth Higgs stratum.  Any
\(B\)-field, Wess--Zumino term, or dilaton/source deformation belongs to the
coupled two-dimensional \((4,4)\) target datum rather than to the pure metric
theorem.
\end{example}

\begin{remark}[D1--D5 throat bridge]
\label{rem:d1-d5-coulomb-throat-higgs-bridge}
The same two-dimensional gauge theory also supplies the QFT bridge that ADHM
alone does not give.  It connects the Higgs quotient to the symmetric product
CFT point.  In a rank-one Coulomb chart, the vector multiplet has scalars
\(\vec X\in\mathbb R^4\), with \(r=|\vec X|\).  The \(Q_5\) charged
hypermultiplets then have masses proportional to \(r\).

One-loop integration of those hypers in the \(\mathcal N=(4,4)\) branch EFT
gives the Coulomb sigma-model datum
\begin{equation}
  ds_C^2=h(r)\,d\vec X^{\,2},
  \qquad
  h(r)=h_\infty+\frac{Q_5}{2r^2},
  \qquad
  H=-\ast_4 dh .
  \label{eq:d1-d5-coulomb-throat-metric-flux}
\end{equation}
The normalization is chosen so that
\begin{equation}
  \Delta_{\mathbb R^4}h=0\quad (r>0),
  \qquad
  \frac{1}{2\pi^2}\int_{S_r^3}H=Q_5 .
  \label{eq:d1-d5-coulomb-hflux-quantization}
\end{equation}
Near \(r=0\), the metric in the logarithmic radial coordinate
\(t=\log r\) has
\[
  h(r)\,r^2\longrightarrow \frac{Q_5}{2},
\]
so the Coulomb branch develops an infinite throat rather than an ordinary
finite-distance endpoint.  This is precisely the kind of branch datum that is
not contained in the hyperk\"ahler quotient
\eqref{eq:d1-d5-adhm-higgs-quotient}: it is a quantum vector-multiplet metric
plus torsion flux, not a protected Higgs metric.

The FI deformation is the gauge-theory mechanism that removes this Coulomb
continuum from the supersymmetric vacuum problem.  Schematically,
\[
  V_\zeta
  =
  \frac{e_2^2}{2}\|\mu_{\mathbb R}-\zeta{\bf1}_K\|^2
  +e_2^2\|\mu_{\mathbb C}\|^2
  +r^2\bigl(\|I\|^2+\|J\|^2\bigr)+\cdots .
\]
At the empty-framing Coulomb locus \(I=J=0\), the trace of the commutator
part of \(\mu_{\mathbb R}\) is zero, while
\(\operatorname{Tr}(\zeta{\bf1}_K)=\zeta Q_1\).  Hence
\begin{equation}
  \|\mu_{\mathbb R}-\zeta{\bf1}_K\|^2
  \ge
  \frac{|\operatorname{Tr}(\mu_{\mathbb R}-\zeta{\bf1}_K)|^2}{Q_1}
  =
  Q_1\zeta^2 ,
  \label{eq:d1-d5-positive-fi-coulomb-lift-bound}
\end{equation}
so a positive FI parameter lifts the empty-framing Coulomb vacuum before any
Higgs-branch quotient metric is interpreted as an infrared CFT metric.

The proposed relation to the symmetric-product point is therefore a bridge
between QFT data, not a dimension-count slogan.  A bridge datum must include
at least:
\[
\begin{gathered}
  \text{the \(\mathcal N=(4,4)\) gauge theory and FI path},\quad
  \text{the protected Higgs metric on the smooth stratum},\\
  \text{torsion/\(B\)-field and flux data},\quad
  \text{the exclusion or decoupling of the Coulomb throat},\\
  \text{an operator map to the symmetric-product blow-up tangent}
  \quad
  \text{of Construction~\ref{constr:n4-symmetric-product-two-cycle-marginal}}.
\end{gathered}
\]
The equality
\(\dim_{\mathbb R}\mathcal M_{Q_1,Q_5}=4Q_1Q_5\) and the symmetric-product
local tangent count of
\eqref{eq:n4-symprod-local-marginal-count} are necessary consistency checks.
They are not sufficient: for example, different pairs \((Q_1,Q_5)\) can have
the same product \(Q_1Q_5\) and hence the same Higgs-branch dimension while
carrying different Coulomb-throat flux \(Q_5\).  The intrinsic D1--D5 QFT
route is the conditional statement that after the positive-FI Higgs stratum is
isolated, the branch sigma model flows to the same compact
\(\mathcal N=(4,4)\) conformal family whose local symmetric-product tangent is
constructed in Volume~V.  The noncompact throat, torsion flux, normalizable
operator map, and global CFT identification group are part of that assertion,
not consequences of the ADHM equations alone.
\end{remark}

\section{Singularities}

A point \(p\in\mathcal M\) is singular when the low-energy description changes
there: additional states can become massless, the effective metric can
degenerate, or the coordinate quotient can have an orbifold or more severe
singularity.  In the \(\mathcal N=2\) pure gauge-algebra \(\mathfrak{su}(2)\)
example, the singularities \(u=\pm\Lambda^2\) are interpreted by vanishing
cycles of the Seiberg-Witten curve and the appearance of light monopole or
dyon hypermultiplets.  The correct local model is the Abelian theory with
those light fields included.

\section{Effective Theory On A Branch}

Let \(p\) lie on a smooth branch and let \(z^i\) be local coordinates.  The
low-energy effective action has the derivative expansion
\[
  S_{\mathrm{eff}}
  =
  \int d^dx\,
  g_{i\bar j}(z,\bar z)\partial_\mu z^i\partial^\mu\bar z^{\bar j}
  +\text{fermions}
  +\text{higher-derivative terms}.
\]
The branch approximation is valid for momenta small compared with the masses
of fields integrated out and away from loci where additional fields become
light.  A complete moduli-space statement therefore includes its domain of
validity inside the full QFT.

\begin{openproblem}[Moduli-space construction]
\label{op:susy-nonperturbative-moduli-space-construction}
Construct the supersymmetric moduli-space datum as part of a general passage
from Wightman-type local QFT data to Kontsevich--Segal-type geometric
amplitude data.  Starting from local operator distributions, Hilbert spaces,
vacuum sectors, and correlation distributions, the construction should
produce the boundary state spaces, sewing maps, reflection pairings, and
background-field dependence from which the vacuum sectors,
chiral-coordinate ring, low-energy Hilbert spaces, branch metrics, singular
local models, and background anomaly responses are recovered.  It should
also explain the precise domain of the two probes described above:
chiral-ring characters as reduced holomorphic-coordinate data, and
fixed-charge spectra as large-charge asymptotic data for the effective
theory on charged regions of the branch.
\end{openproblem}
