\chapter{All-Loop Asymptotic Bethe Ansatz}
\label{ch:planar-n4-asymptotic-bethe-ansatz}

\ConventionsLorentzian

The one-loop spin chain of
Chapter~\ref{ch:planar-n4-spectral-problem-spin-chains} is only the first
term in the planar dilatation operator.  The all-loop asymptotic Bethe ansatz
is a proposal for the spectrum of long single-trace operators before wrapping
effects become important.  Its inputs are the residual centrally extended
\(\mathfrak{su}(2|2)\oplus\mathfrak{su}(2|2)\) symmetry of the BMN vacuum,
factorized magnon scattering, crossing, and analyticity.

\section{Convention Ledger}

The integrability literature uses several mutually inverse scalar-factor and
Zhukovsky conventions.  This chapter follows the convention used in the
stringbook integrability chapters, with the notational replacement
\[
  h_{\rm stringbook}=g_{\rm monograph}=\frac{\sqrt\lambda}{4\pi}.
\]
Thus
\[
  x^\pm+\frac1{x^\pm}
  =
  \frac{u\pm\ii/2}{g},
  \qquad
  \ee^{\ii p}=\frac{x^+}{x^-},
\]
and the physical weak-coupling branch has
\[
  x^\pm
  =
  \frac{\ee^{\pm\ii p/2}}{2g\sin(p/2)}+O(g).
\]
The one-magnon energy is written in the stringbook-compatible form
\[
  H
  =
  -1-2\ii g(x^+-x^-)
  =
  1+2\ii g\left(\frac1{x^+}-\frac1{x^-}\right).
\]
The equality of the two expressions is not an additional assumption: it is
obtained by multiplying the defining \(x^\pm\) relation by \(2\ii g\).

The scalar factor convention is also fixed here.  We write
\(\sigma=\ee^{\ii\theta}\) and use the crossing equation in the same
orientation as the stringbook:
\[
  \sigma_{12}\sigma_{\bar1 2}
  =
  \frac{x_2^+}{x_2^-}
  \frac{x_2^+ -x_1^+}{x_2^+ -x_1^-}
  \frac{x_2^- -1/x_1^+}{x_2^- -1/x_1^-}.
\]
Authors who put the rational scalar factor into the opposite part of the
S-matrix, or define the two-body operator as \(S_{21}\) rather than \(S_{12}\),
write the reciprocal of this expression.  The companion check script enforces
the \(x^\pm\)-energy and sheet conventions used here; crossing itself remains
a monodromy input rather than a finite algebraic identity.

\begin{warning}[Status of magnon crossing]
\label{warn:planar-n4-magnon-crossing-status}
The crossing equation used in this chapter is not the Wightman--LSZ crossing
theorem for four-dimensional local scattering amplitudes.  A spin-chain
magnon is an excitation above the chosen BMN vacuum in a single-trace sector;
it is not created by a gauge-invariant local field, its two-body scattering
operator is asymptotic in the trace length, and its dispersion is encoded on
the Zhukovsky rapidity surface rather than on a relativistic mass shell.
Consequently ordinary four-dimensional QFT does not by itself give the
replacement \(x^\pm\mapsto1/x^\pm\) or the scalar equation displayed above.

The conceptual origin of the condition is the two-dimensional relativistic
integrable sigma-model description that appears in the strong-coupling
AdS/CFT analysis.  Before uniform light-cone gauge fixing, that auxiliary
two-dimensional theory has a Euclidean continuation in which mirror-channel
interchange of space and time and relativistic crossing are realized as
analytic continuation in rapidity.  Gauge fixing and decompactification turn
the excitations into nonrelativistic magnons with the Zhukovsky dispersion,
so the inherited crossing operation has no intrinsic explanation from the
nonrelativistic spin chain alone.  In the resulting integrable
sigma-model/spin-chain framework the crossed magnon is the contragredient
short representation of the centrally extended residual symmetry;
equivalently, the antipode operation sends the Zhukovsky variables to the
crossed sheet \(x^\pm\mapsto1/x^\pm\) along a specified path.  Matrix
consistency with this crossed representation fixes the rational multiplier
below, while the dressing phase is required to have the corresponding scalar
monodromy.  Extending this strong-coupling crossing condition to all 't Hooft
coupling, and using it as a statement about planar gauge-theory anomalous
dimensions, is an integrability assumption constrained by symmetry,
unitarity, Yang--Baxter factorization, perturbative gauge-theory checks, and
finite-size TBA/QSC tests.  It is not presently a theorem derived from a
constructive four-dimensional definition of \(\mathcal N=4\) SYM.
\end{warning}

\section{BMN Vacuum and Central Extension}

The vacuum spin-chain state is
\[
  \operatorname{tr}(Z^J).
\]
It is protected and carries large \(R\)-charge \(J\).  A magnon is an
insertion into this long chain.  The residual symmetry preserving the vacuum
contains two centrally extended \(\mathfrak{su}(2|2)\) algebras.  For one
copy, write the bosonic states as \(\ket{\phi^a}\), \(a=1,2\), and the
fermionic states as \(\ket{\psi^\alpha}\), \(\alpha=1,2\).  The supercharges
act in the form
\[
\begin{aligned}
  Q^\alpha{}_a\ket{\phi^b}
  &=a\,\delta_a^b\ket{\psi^\alpha},
&
  Q^\alpha{}_a\ket{\psi^\beta}
  &=b\,\varepsilon^{\alpha\beta}\varepsilon_{ab}\ket{\phi^b Z^+},
\\
  S^a{}_\alpha\ket{\phi^b}
  &=c\,\varepsilon^{ab}\varepsilon_{\alpha\beta}\ket{\psi^\beta Z^-},
&
  S^a{}_\alpha\ket{\psi^\beta}
  &=d\,\delta_\alpha^\beta\ket{\phi^a}.
\end{aligned}
\]
The symbols \(Z^+\) and \(Z^-\) record insertion or removal of one vacuum
letter.  On a momentum eigenstate,
\[
  Z^+\mapsto \ee^{\ii p}-1,
  \qquad
  Z^-\mapsto \ee^{-\ii p}-1
\]
up to the chosen side convention.  The representation constraint is
\[
  ad-bc=1.
\]
The central charges are
\[
  P=ab,\qquad K=cd,\qquad H=ad+bc .
\]
The shortening condition of the centrally extended algebra is
\[
  H^2-4PK=1 .
\]
The central extension is compatible with trace cyclicity.  A single magnon is
an excitation on an open segment of the BMN vacuum, so \(Z^\pm\) records
length data at the edge of that segment.  On a closed trace the product of
all such phases is \(\ee^{\ii P_{\rm tot}}-1\), where \(P_{\rm tot}\) is the
total spin-chain momentum.  The physical single-trace projection sets
\(P_{\rm tot}=0\), but individual magnons may still transform in centrally
extended short representations before that projection is imposed.
With
\[
  P=g(1-\ee^{\ii p}),\qquad K=g(1-\ee^{-\ii p}),
\]
this gives the exact single-magnon dispersion relation
\begin{equation}
  H(p)=\sqrt{1+16g^2\sin^2\frac p2}.
  \label{eq:planar-n4-magnon-dispersion}
\end{equation}
The anomalous contribution of a magnon is \(H(p)-1\).

\begin{controlledapproximation}[BMN scaling check]
\label{ca:planar-n4-bmn-scaling-check}
Let \(J\to\infty\) with \(\lambda'=\lambda/J^2\) fixed and take
\[
  p=\frac{2\pi n}{J}+O(J^{-2}),\qquad n\in\mathbb Z .
\]
Then
\[
  H(p)=\sqrt{1+\lambda' n^2}+O(J^{-1}).
\]
This is the all-coupling continuation of the gauge-theory BMN normalization
derived at one loop in
Proposition~\ref{prop:planar-n4-two-magnon-bmn-quantization}.  Here it is a
convention check on the spectral problem, not an external derivation of the
all-coupling Bethe ansatz.
\end{controlledapproximation}

\paragraph{Dispersion from shortening.}
Assuming the centrally extended residual symmetry acts on a one-magnon
representation with central charges \(P=g(1-\ee^{\ii p})\) and
\(K=g(1-\ee^{-\ii p})\), the shortening condition gives the dispersion
relation by the finite algebraic check
\[
  PK
  =
  g^2(1-\ee^{\ii p})(1-\ee^{-\ii p})
  =
  4g^2\sin^2\frac p2 .
\]
Substitution into \(H^2-4PK=1\) gives
\[
  H^2=1+16g^2\sin^2\frac p2 .
\]
The positive square root is selected because \(H=1\) at \(g=0\) for an
elementary magnon.

\section{Zhukovsky Variables}

Introduce the Zhukovsky map \(x(u)\) by
\begin{equation}
  x(u)+\frac1{x(u)}=\frac ug .
  \label{eq:planar-n4-zhukovsky-map}
\end{equation}
The physical branch is the solution satisfying \(x(u)\sim u/g\) as
\(|u|\to\infty\).  Equivalently,
\[
  x_{\rm ph}(u)
  =
  \frac12\left(
    \frac ug+\sqrt{\frac ug-2}\sqrt{\frac ug+2}
  \right),
\]
with the square-root cuts chosen so that the only short cut on the physical
sheet is \([-2g,2g]\).  Crossing this cut replaces \(x\) by \(1/x\).  More
precisely, the two boundary values on the cut satisfy
\[
  x_{\rm above}(u)x_{\rm below}(u)=1,
  \qquad -2g<u<2g .
\]
The large-\(u\) expansion on the physical sheet is
\[
  x_{\rm ph}(u)
  =
  \frac ug-\frac gu-\frac{g^3}{u^3}
  -\frac{2g^5}{u^5}+O(u^{-7}).
\]
For a one-magnon state set
\[
  x^\pm=x(u\pm\ii/2).
\]
Then
\begin{equation}
  \ee^{\ii p}=\frac{x^+}{x^-},
  \qquad
  x^++\frac1{x^+}-x^- -\frac1{x^-}=\frac{\ii}{g}.
  \label{eq:planar-n4-xpm-definition}
\end{equation}
The physical region for real magnons is chosen so that
\((x^+)^*=x^-\) and \(|x^\pm|>1\) at weak coupling.  The magnon energy is
the sheet-dependent but single-valued physical-branch expression
\begin{equation}
  H
  =
  -1-2\ii g(x^+-x^-)
  =
  1+2\ii g\left(
    \frac1{x^+}-\frac1{x^-}
  \right),
  \label{eq:planar-n4-energy-xpm}
\end{equation}
with the same branch as \eqref{eq:planar-n4-magnon-dispersion}.  Conserved
charges used in the dressing phase are
\[
  q_r(x^\pm)
  =
  \frac{\ii}{r-1}
  \left[
    \frac1{(x^+)^{r-1}}-\frac1{(x^-)^{r-1}}
  \right],
  \qquad r\geq2 .
\]

\paragraph{Zhukovsky form of the dispersion.}
Let \(x^\pm\) obey \eqref{eq:planar-n4-xpm-definition} on the physical sheet.
Then \eqref{eq:planar-n4-energy-xpm} is equal to
\[
  \sqrt{1+16g^2\sin^2\frac p2}.
\]

For a physical real magnon write
\[
  x^\pm
  =
  \ee^{\pm\ii p/2}
  \frac{1+\sqrt{1+16g^2\sin^2(p/2)}}
       {4g\sin(p/2)} .
\]
This parametrization gives \(x^+/x^-=\ee^{\ii p}\).  Substituting it into the
difference \(x^++1/x^+-x^- -1/x^-\) gives \(\ii/g\), so it is the physical
solution of \eqref{eq:planar-n4-xpm-definition}.  The inverse difference is
\[
  \frac1{x^+}-\frac1{x^-}
  =
  \frac{\sqrt{1+16g^2\sin^2(p/2)}-1}{2\ii g}.
\]
Equation~\eqref{eq:planar-n4-energy-xpm} therefore reproduces
\eqref{eq:planar-n4-magnon-dispersion}.  The equality extends to the
physical sheet by analytic continuation away from the branch points.

\begin{warning}[Sheet data in crossing]
\label{warn:planar-n4-crossing-sheet-data}
The formal replacement \(x^\pm\mapsto1/x^\pm\) is not a harmless algebraic
substitution.  It records a path on the Zhukovsky surface.  Poles crossed by
that path alter the scalar factor through the crossing equation, and the
answer depends on whether the continuation crosses the short cut of \(x^+\),
the short cut of \(x^-\), or the composed path used in the mirror theory.
In the \(u\)-plane, the cuts of \(x^\pm=x(u\pm\ii/2)\) have endpoints
\[
  u=\pm2g-\ii/2,\qquad u=\pm2g+\ii/2 .
\]
The physical crossing path for one magnon winds around the appropriate pair
of these branch points so that both \(x^+\) and \(x^-\) reach the crossed
sheet.  Applying the path twice generally lands on a different sheet of the
scalar factor even though the projected \(x^\pm\)-values have returned.
\end{warning}

\begin{definition}[Crossing path for the Zhukovsky variables]
\label{def:planar-n4-crossing-path}
Let \(x(u)\) be the physical Zhukovsky branch, \(x(u)\sim u/g\) at infinity,
and set \(x^\pm(u)=x(u\pm\ii/2)\).  A crossing path for the first magnon is
a path \(\gamma_{\rm cr}\) on the \(u\)-surface, starting at a physical point
and ending above the same projected \(u\)-value on a crossed sheet, with the
following topological data:
\[
  w_{\gamma_{\rm cr}}\!\left(2g-\frac{\ii}{2}\right)=1,
  \qquad
  w_{\gamma_{\rm cr}}\!\left(2g+\frac{\ii}{2}\right)=1
  \pmod 2,
\]
and with even winding around the other two shifted endpoints in this
right-endpoint convention.  Equivalently, the lower shifted cut of \(x^+\)
and the upper shifted cut of \(x^-\) are each crossed once.  When this path
is used in the DHM contour formula below, it is also required to avoid the
Cauchy and Gamma-function singularities except for the controlled residue
crossings described in
Proposition~\ref{prop:planar-n4-dhm-local-residue-continuation}.
\end{definition}

\paragraph{Zhukovsky monodromy along the crossing path.}
Analytic continuation of \(x^\pm(u)\) along a path with the winding data of
Definition~\ref{def:planar-n4-crossing-path} gives
\[
  \bar x^\pm(u)=\frac{1}{x^\pm(u)} .
\]
Consequently,
\[
  \frac{\bar x^+}{\bar x^-}
  =
  \left(\frac{x^+}{x^-}\right)^{-1},
  \qquad
  H(\bar x^\pm)=-H(x^\pm),
\]
where \(H(x^\pm)=1+2\ii g(1/x^+-1/x^-)\) is the physical-sheet energy
continued as a meromorphic function of \(x^\pm\).

Write
\[
  x(u)=\frac12\left(\frac{u}{g}
  +\sqrt{\left(\frac{u}{g}\right)^2-4}\right)
\]
on the physical branch.  Encircling either endpoint of the short cut
\([-2g,2g]\) changes the sign of the square root and therefore sends
\[
  x(u)\longmapsto
  \frac12\left(\frac{u}{g}
  -\sqrt{\left(\frac{u}{g}\right)^2-4}\right)
  =
  \frac1{x(u)} .
\]
For \(x^+(u)=x(u+\ii/2)\), the relevant shifted cut has endpoint
\(2g-\ii/2\); for \(x^-(u)=x(u-\ii/2)\), it has endpoint \(2g+\ii/2\).
The mod-two winding data in Definition~\ref{def:planar-n4-crossing-path}
therefore inverts both \(x^+\) and \(x^-\).  The momentum ratio is then
inverted immediately.  Finally, using the stringbook energy identity
\[
  H(x^\pm)=-1-2\ii g(x^+-x^-)
  =
  1+2\ii g(1/x^+-1/x^-)
\]
on the physical sheet, the crossed value is
\[
  H(\bar x^\pm)
  =
  1+2\ii g(x^+-x^-)
  =
  -H(x^\pm).
\]

\section{Two-Body S-Matrix}

The magnon two-body \(S\)-matrix is fixed by symmetry up to a scalar factor,
after one has fixed the dynamic spin-chain frame.  Let
\(S^{\rm mat}_{12}\) denote the matrix part determined by the two centrally
extended \(\mathfrak{su}(2|2)\) copies.  The full two-magnon operator is
\[
  S_{12}=S_0(x_1^\pm,x_2^\pm)\,S^{\rm mat}_{12}.
\]
For one \(\mathfrak{su}(2|2)\) copy, write \(S^{(1)}_{12}\) for the
intertwiner before tensoring the left and right copies.  It is defined by
\[
  S^{(1)}_{12}\Delta_{12}(J)
  =
  \Delta_{21}(J)S^{(1)}_{12},
  \qquad J\in\mathfrak{su}(2|2)_{\rm c},
\]
where the coproduct includes the braiding phase carried by the magnon momentum.
The labels \(1,2\) on one-particle states record the momenta \(p_1,p_2\), not
extra internal indices.  We use unnormalized symmetrization and
antisymmetrization:
\[
  v^{\{a}w^{b\}}=v^a w^b+v^b w^a,
  \qquad
  v^{[a}w^{b]}=v^a w^b-v^b w^a .
\]
With this convention the most general
\(\mathfrak{su}(2)\oplus\mathfrak{su}(2)\)-covariant two-particle operator has
ten amplitudes:
\[
\begin{aligned}
  S^{(1)}_{12}\ket{\phi_1^a\phi_2^b}
  &=
  A_{12}\ket{\phi_2^{\{a}\phi_1^{b\}}}
  +B_{12}\ket{\phi_2^{[a}\phi_1^{b]}}
  +\frac12 C_{12}\varepsilon^{ab}\varepsilon_{\alpha\beta}
    \ket{\psi_2^\alpha\psi_1^\beta},\\
  S^{(1)}_{12}\ket{\psi_1^\alpha\psi_2^\beta}
  &=
  D_{12}\ket{\psi_2^{\{\alpha}\psi_1^{\beta\}}}
  +E_{12}\ket{\psi_2^{[\alpha}\psi_1^{\beta]}}
  +\frac12 F_{12}\varepsilon^{\alpha\beta}\varepsilon_{ab}
    \ket{\phi_2^a\phi_1^b},\\
  S^{(1)}_{12}\ket{\phi_1^a\psi_2^\beta}
  &=
  G_{12}\ket{\psi_2^\beta\phi_1^a}
  +H_{12}\ket{\phi_2^a\psi_1^\beta},
  \\
  S^{(1)}_{12}\ket{\psi_1^\alpha\phi_2^b}
  &=
  K_{12}\ket{\psi_2^\alpha\phi_1^b}
  +L_{12}\ket{\phi_2^b\psi_1^\alpha}.
\end{aligned}
\]
The length-changing markers are suppressed in this display but not forgotten:
the \(C_{12}\) channel carries the \(Z^-\) marker and the \(F_{12}\) channel
carries the \(Z^+\) marker in the stringbook dynamic spin-chain frame.  The
markers are precisely the source of the braided coproduct.

For a magnon \(i\) write its representation coefficients as
\(a_i,b_i,c_i,d_i\).  The stringbook frame relates them to \(x_i^\pm\) by
\[
  x_i^+=\ii\frac{d_i}{b_i},
  \qquad
  x_i^-=-\ii\frac{a_i}{c_i},
  \qquad
  a_i d_i-b_i c_i=1 .
\]
The same \(x_i^\pm\) obey the Zhukovsky relations already fixed in
\eqref{eq:planar-n4-xpm-definition}.  At generic non-pole kinematics the
tensor product of two fundamental short representations is long and the
intertwiner is unique up to one scalar.  Let this one-copy scalar be denoted
\(\mathcal S_{12}\).  Solving the finite intertwining equations gives
\begin{equation}
\begin{aligned}
  A_{12}
  &=
  \mathcal S_{12}
  \frac{x_2^+-x_1^-}{x_2^- -x_1^+},
  \\
  B_{12}
  &=
  A_{12}\left[
  1
  -2\frac{x_2^+-x_1^+}{x_2^+-x_1^-}
     \frac{x_2^- -1/x_1^+}{x_2^- -1/x_1^-}
  \right],
  \\
  C_{12}
  &=
  \mathcal S_{12}
  \frac{2g^{-1}a_1a_2}{x_1^-x_2^- -1}
  \frac{x_2^+-x_1^+}{x_2^- -x_1^+},
  \\
  D_{12}
  &=-\mathcal S_{12},
  \\
  E_{12}
  &=
  -\mathcal S_{12}\left[
  1
  -2\frac{x_2^- -x_1^-}{x_2^- -x_1^+}
     \frac{x_2^+ -1/x_1^-}{x_2^+ -1/x_1^+}
  \right],
  \\
  F_{12}
  &=
  -\mathcal S_{12}
  \frac{2g(x_1^+-x_1^-)(x_2^+-x_2^-)}
       {a_1a_2(x_1^+x_2^+ -1)}
  \frac{x_2^- -x_1^-}{x_2^- -x_1^+},
  \\
  G_{12}
  &=
  \mathcal S_{12}
  \frac{x_2^+-x_1^+}{x_2^- -x_1^+},
  \qquad
  H_{12}
  =
  \mathcal S_{12}\frac{a_1}{a_2}
  \frac{x_2^+-x_2^-}{x_2^- -x_1^+},
  \\
  K_{12}
  &=
  \mathcal S_{12}\frac{a_2}{a_1}
  \frac{x_1^+-x_1^-}{x_2^- -x_1^+},
  \qquad
  L_{12}
  =
  \mathcal S_{12}
  \frac{x_2^- -x_1^-}{x_2^- -x_1^+}.
\end{aligned}
\label{eq:planar-n4-su2c-ten-amplitudes}
\end{equation}

\begin{proposition}[Generic rank certificate for the one-copy intertwiner]
\label{prop:planar-n4-su2c-intertwiner-rank-certificate}
Let
\[
  d_{12}=x_2^- -x_1^+,\qquad n_{12}=x_2^+ -x_1^-,
\]
and work on the generic open set where
\[
  x_1^\pm x_2^\pm d_{12}n_{12}a_1a_2
  (x_1^-x_2^- -1)(x_1^+x_2^+ -1)\ne0 .
\]
After the dynamic-frame \(Q\)- and \(S\)-intertwining equations are written
on the ten \(\mathfrak{su}(2)\oplus\mathfrak{su}(2)\)-covariant amplitudes
and the representation ratios
\[
  x_i^+=\ii d_i/b_i,\qquad x_i^-=-\ii a_i/c_i,\qquad a_id_i-b_ic_i=1
\]
are used, the finite linear system has the following row-echelon chart:
\begin{equation}
\begin{gathered}
  B_{12}=\beta_{12}A_{12},\qquad
  C_{12}=\gamma_{12}A_{12},\qquad
  D_{12}=\delta_{12}A_{12},\qquad
  E_{12}=\varepsilon_{12}A_{12},\\
  F_{12}=\varphi_{12}A_{12},\qquad
  G_{12}=g_{12}A_{12},\qquad
  H_{12}=h_{12}A_{12},\qquad
  K_{12}=k_{12}A_{12},\qquad
  L_{12}=\ell_{12}A_{12},
\end{gathered}
\label{eq:planar-n4-su2c-row-chart}
\end{equation}
where
\begin{equation}
\begin{aligned}
  \beta_{12}
  &=
  1
  -2\frac{x_2^+-x_1^+}{n_{12}}
     \frac{x_2^- -1/x_1^+}{x_2^- -1/x_1^-},
  &
  \gamma_{12}
  &=
  \frac{2a_1a_2}{g}
  \frac{x_2^+-x_1^+}{(x_1^-x_2^- -1)n_{12}},
  \\
  \delta_{12}
  &=-\frac{d_{12}}{n_{12}},
  &
  \varepsilon_{12}
  &=
  -\frac{d_{12}}{n_{12}}
  \left[
  1
  -2\frac{x_2^- -x_1^-}{d_{12}}
     \frac{x_2^+ -1/x_1^-}{x_2^+ -1/x_1^+}
  \right],
  \\
  \varphi_{12}
  &=
  -\frac{2g}{a_1a_2}
  \frac{(x_1^+-x_1^-)(x_2^+-x_2^-)(x_2^- -x_1^-)}
       {(x_1^+x_2^+ -1)n_{12}},
  &
  g_{12}
  &=
  \frac{x_2^+-x_1^+}{n_{12}},
  \\
  h_{12}
  &=
  \frac{a_1}{a_2}\frac{x_2^+-x_2^-}{n_{12}},
  &
  k_{12}
  &=
  \frac{a_2}{a_1}\frac{x_1^+-x_1^-}{n_{12}},
  \\
  \ell_{12}
  &=
  \frac{x_2^- -x_1^-}{n_{12}} .
\end{aligned}
\label{eq:planar-n4-su2c-row-coefficients}
\end{equation}
Consequently the solution space of the finite intertwining equations is
one-dimensional on this open set.  Choosing
\[
  A_{12}=\mathcal S_{12}\frac{n_{12}}{d_{12}}
\]
recovers the ten amplitudes in
\eqref{eq:planar-n4-su2c-ten-amplitudes}.
\end{proposition}

\begin{proof}
The displayed chart is a row-reduced certificate.  The unknown vector is
\[
  (A,B,C,D,E,F,G,H,K,L).
\]
Each equation in \eqref{eq:planar-n4-su2c-row-chart} has the form
\[
  X-r_XA=0,\qquad X\in\{B,C,D,E,F,G,H,K,L\},
\]
with \(r_X\) one of the nine coefficients in
\eqref{eq:planar-n4-su2c-row-coefficients}.  Ordered by
\((B,C,D,E,F,G,H,K,L)\), these nine equations have pivot coefficient \(1\)
in nine distinct columns.  Their coefficient matrix therefore has rank
nine, so its kernel in the ten-dimensional amplitude space is one
dimensional and is parametrized by \(A\).

It remains only to identify the parametrization with the convention used in
\eqref{eq:planar-n4-su2c-ten-amplitudes}.  Substituting
\(A_{12}=\mathcal S_{12}n_{12}/d_{12}\) into
\eqref{eq:planar-n4-su2c-row-chart} gives, term by term, the formulas for
\(B,C,D,E,F,G,H,K,L\) in
\eqref{eq:planar-n4-su2c-ten-amplitudes}.  The genericity assumptions are
exactly the denominators that appear in this chart and in the two channels
where the \(Z^\pm\) markers enter.  At their zero loci the tensor product can
become reducible or the chosen dynamic-frame chart can degenerate; those
special loci are the bound-state and pole analyses, not the generic
two-magnon scattering case.
\end{proof}

\paragraph{Local intertwining and block-unitarity checks.}
\label{par:planar-n4-su2c-local-amplitude-checks}
Assume the denominators in
\eqref{eq:planar-n4-su2c-ten-amplitudes} are non-zero.  The amplitudes above
satisfy the highest-weight \(Q\)-intertwining relation
\begin{equation}
  A_{12}
  =
  \frac{a_1}{a_2}K_{12}+G_{12}
  =
  L_{12}+\frac{a_2}{a_1}H_{12}.
  \label{eq:planar-n4-su2c-q-check}
\end{equation}
After factoring out \(\mathcal S_{12}\), the same amplitudes also obey the
local analytic unitarity identities
\[
  R_{12}^{\rm BB}R_{21}^{\rm BB}=1,
  \qquad
  R_{12}^{\rm BF}R_{21}^{\rm BF}=1,
\]
on the boson-boson/fermion-fermion block \(R^{\rm BB}\) and the mixed
boson-fermion block \(R^{\rm BF}\) written, for fixed distinct doublet
indices, in the ordered bases
\[
  (\psi_1^1\psi_2^2,\psi_1^2\psi_2^1,\phi_1^1\phi_2^2,\phi_1^2\phi_2^1),
  \qquad
  (\phi_1^1\psi_2^1,\psi_1^1\phi_2^1).
\]
The blocks are explicitly
\[
  R_{12}^{\rm BB}
  =
  \frac1{\mathcal S_{12}}
  \begin{pmatrix}
    (D-E)/2 & (D+E)/2 & -F/2 & F/2\\
    (D+E)/2 & (D-E)/2 & F/2 & -F/2\\
    -C/2 & C/2 & (A-B)/2 & (A+B)/2\\
    C/2 & -C/2 & (A+B)/2 & (A-B)/2
  \end{pmatrix}_{12},
  \qquad
  R_{12}^{\rm BF}
  =
  \frac1{\mathcal S_{12}}
  \begin{pmatrix}
    H & G\\
    L & K
  \end{pmatrix}_{12}.
\]
Equivalently, the matrix part has the correct inverse when the two magnons are
exchanged; the scalar unitarity condition is the separate condition
\(\mathcal S_{12}\mathcal S_{21}=1\), or, for the full two-copy scalar split,
\(S_{0,12}S_{0,21}=1\).

The finite check is local in the chosen dynamic frame.  The first identity is
the result of applying \(Q^\alpha{}_a\) to the symmetric boson state and
comparing the independent
\(\ket{\psi_2\phi_1}\) and \(\ket{\phi_2\psi_1}\) components.  Substituting
\eqref{eq:planar-n4-su2c-ten-amplitudes} gives
\[
\begin{aligned}
  \frac{a_1}{a_2}K_{12}+G_{12}
  &=
  \mathcal S_{12}
  \frac{x_1^+-x_1^-+x_2^+-x_1^+}{x_2^- -x_1^+}
  =
  \mathcal S_{12}
  \frac{x_2^+-x_1^-}{x_2^- -x_1^+}
  =
  A_{12},\\
  L_{12}+\frac{a_2}{a_1}H_{12}
  &=
  \mathcal S_{12}
  \frac{x_2^- -x_1^-+x_2^+-x_2^-}{x_2^- -x_1^+}
  =
  A_{12}.
\end{aligned}
\]
The remaining \(Q\)- and \(S\)-intertwining equations are obtained in the same
finite way from the other highest-weight components.  Clearing denominators
reduces them to polynomial identities in \(x_i^\pm\) and the representation
ratios \(a_1/a_2\), together with the defining relations for \(b_i,c_i,d_i\)
above.  The companion calculation check constructs the two nontrivial finite
blocks explicitly from \eqref{eq:planar-n4-su2c-ten-amplitudes} and verifies
\(R_{12}R_{21}=1\) on nonsingular physical-branch samples.  This check is not
a substitute for crossing or Yang--Baxter factorization; it certifies the
local frame and amplitude algebra used by the nested Bethe derivation below.

The convenient scalar split used in the stringbook is
\[
  \mathcal S_{12}
  =
  \left[
  \frac{x_2^- -x_1^+}{x_2^+ -x_1^-}
  \frac{1-1/(x_2^+x_1^-)}{1-1/(x_2^-x_1^+)}
  \right]^{1/2}\sigma_{12}.
\]
Consequently the compact \(SU(2)\)-sector two-copy amplitude
\((A_{12})^2\) contains the rational factor
\[
  \frac{x_2^+-x_1^-}{x_2^- -x_1^+}
  \frac{1-1/(x_2^+x_1^-)}{1-1/(x_2^-x_1^+)}
\]
times \(\sigma_{12}^2\), in the orientation fixed by the convention ledger.
Equivalently, the full scalar factor may be written as
\[
  S_0(x_1^\pm,x_2^\pm)
  =
  S_{\rm rat}(x_1^\pm,x_2^\pm)\,
  \sigma(x_1^\pm,x_2^\pm)^2,
  \qquad
  \sigma=\ee^{\ii\theta}.
\]
The rational part is fixed by the same representation data that determine
the matrix part.  The dressing phase has the antisymmetric charge expansion
\[
  \theta(x_1^\pm,x_2^\pm)
  =
  -
  \sum_{r=2}^\infty\sum_{\substack{s>r\\ r+s\ {\rm odd}}}
  c_{r,s}(g)\,
  \bigl(q_r(x_1^\pm)q_s(x_2^\pm)-q_s(x_1^\pm)q_r(x_2^\pm)\bigr).
\]

\paragraph{Dressing unitarity from charge antisymmetry.}
Work in a domain of the physical sheet where the charge expansion for
\(\theta_{12}=\theta(x_1^\pm,x_2^\pm)\) is either finite after truncation or
absolutely convergent, and use the same branch of the phase for the exchanged
pair \(\theta_{21}=\theta(x_2^\pm,x_1^\pm)\).  Then
\[
  \theta_{21}=-\theta_{12},
  \qquad
  \sigma_{12}\sigma_{21}=1.
\]
Consequently the squared dressing factor appearing in the \(SU(2)\)
asymptotic Bethe equation also satisfies
\[
  \sigma_{12}^2\sigma_{21}^2=1 .
\]

For each pair \(r<s\) with \(r+s\) odd, define
\[
  A_{r,s}(1,2)
  =
  q_r(x_1^\pm)q_s(x_2^\pm)-q_s(x_1^\pm)q_r(x_2^\pm).
\]
Exchanging the two magnons gives
\[
  A_{r,s}(2,1)
  =
  q_r(x_2^\pm)q_s(x_1^\pm)-q_s(x_2^\pm)q_r(x_1^\pm)
  =
  -A_{r,s}(1,2).
\]
The coefficients \(c_{r,s}(g)\) are scalar functions of the coupling and the
integer pair \((r,s)\), not of the ordering of the two magnons.  Therefore the
whole series changes sign under \(1\leftrightarrow2\), proving
\(\theta_{21}=-\theta_{12}\).  Since \(\sigma=\ee^{\ii\theta}\),
\[
  \sigma_{12}\sigma_{21}
  =
  \ee^{\ii(\theta_{12}+\theta_{21})}
  =
  1.
\]
Squaring gives the last statement.

This is only the scalar dressing part of physical two-body unitarity.  The
rational scalar factor and the matrix part must obey their own unitarity
relations in the same frame.  The computation also does not replace crossing:
after analytic continuation around Zhukovsky cuts the charge expansion must be
continued with its prescribed monodromy, and that monodromy is precisely the
additional crossing datum discussed below.

Equivalently, the BES phase may be packaged as
\begin{equation}
\begin{aligned}
  \theta(x_1^\pm,x_2^\pm)
  &=
  \chi(x_1^+,x_2^+)-\chi(x_1^+,x_2^-)
  -\chi(x_1^-,x_2^+)+\chi(x_1^-,x_2^-),\\
  \chi(x,y)
  &=
  -\ii
  \oint\frac{\dd z}{2\pi\ii}
  \oint\frac{\dd w}{2\pi\ii}
  \frac{1}{x-z}\frac{1}{y-w}
  \log
  \frac{\Gamma\!\left(1+\ii g(z+z^{-1}-w-w^{-1})\right)}
       {\Gamma\!\left(1-\ii g(z+z^{-1}-w-w^{-1})\right)} .
\end{aligned}
  \label{eq:planar-n4-bes-chi-integral}
\end{equation}
The contours encircle the unit circle singularities of \(z\) and \(w\) in the
domain where \(|x|>1\), \(|y|>1\).  Formula
\eqref{eq:planar-n4-bes-chi-integral} is therefore not a global formula on
all sheets by itself; outside that domain it is continued with the crossing
prescription.

\begin{proposition}[Local DHM residue continuation]
\label{prop:planar-n4-dhm-local-residue-continuation}
Let
\[
  \mathcal L_g(z,w)
  =
  \log
  \frac{\Gamma\!\left(1+\ii g(z+z^{-1}-w-w^{-1})\right)}
       {\Gamma\!\left(1-\ii g(z+z^{-1}-w-w^{-1})\right)} .
\]
Fix \(x,y\) away from the unit circle and away from the Gamma-function
poles.  Denote by \(\chi_{\rm raw}^{\epsilon_x,\epsilon_y}(x,y)\) the same
unit-circle double contour integral as in
\eqref{eq:planar-n4-bes-chi-integral}, with the \(x\)-Cauchy kernel expanded
radially outside the contour when \(\epsilon_x=+\) and inside the contour
when \(\epsilon_x=-\), and similarly for \(y\).  Thus
\(\chi_{\rm raw}^{+,+}\) is the original DHM branch.  With the stringbook
orientation for crossing inward through \(|x|=1\) and \(|y|=1\), define the
one-residue corrections by
\begin{equation}
  \Psi_x^{\epsilon_y}(x,y)
  =
  \chi_{\rm raw}^{-,\epsilon_y}(x,y)
  -
  \chi_{\rm raw}^{+,\epsilon_y}(x,y),
  \qquad
  \Psi_y^{\epsilon_x}(x,y)
  =
  \chi_{\rm raw}^{\epsilon_x,+}(x,y)
  -
  \chi_{\rm raw}^{\epsilon_x,-}(x,y).
  \label{eq:planar-n4-dhm-single-residue-definitions}
\end{equation}
Equivalently, \(\Psi_x\) is the residue picked up by the pole \(z=x\), with
the sign fixed by the kernel \(1/(x-z)\), while \(\Psi_y\) is written with
the opposite sign because the stringbook formula adds the \(y\)-crossing
correction.  The analytic continuation of the original outside branch into
the three neighboring radial chambers is
\begin{align}
  \chi^{-,+}_{\rm cont}(x,y)
  &=
  \chi_{\rm raw}^{-,+}(x,y)-\Psi_x^{+}(x,y), \\
  \chi^{+,-}_{\rm cont}(x,y)
  &=
  \chi_{\rm raw}^{+,-}(x,y)+\Psi_y^{+}(x,y), \\
  \chi^{-,-}_{\rm cont}(x,y)
  &=
  \chi_{\rm raw}^{-,-}(x,y)
  -\Psi_x^{-}(x,y)
  +\Psi_y^{-}(x,y)
  -\mathcal C_g(x,y),
  \label{eq:planar-n4-dhm-double-residue-continuation}
\end{align}
where the contact term is the double residue
\begin{equation}
  \mathcal C_g(x,y)=\ii\,\mathcal L_g(x,y).
  \label{eq:planar-n4-dhm-contact-term}
\end{equation}
These formulas are local sheet-transition rules for the DHM contour.  The
global Janik crossing equation additionally specifies a path on the
\(u\)-surface, hence which of \(x^\pm(u)\) winds around which shifted
Zhukovsky branch point.
\end{proposition}

\begin{proof}
The only singularities that move when \(x\) crosses the unit circle are the
Cauchy poles \(z=x\); the Gamma-function poles are excluded by hypothesis.
For fixed \(y\)-chamber, deforming the \(z\)-contour back to the unit circle
therefore changes the raw integral exactly by the residue at \(z=x\).  This
gives the first definition in
\eqref{eq:planar-n4-dhm-single-residue-definitions} and hence the first
continuation formula.  The \(y\)-crossing computation is the same, except
that the sign is recorded in \(\Psi_y\) so that the stringbook notebook
convention reads as addition in the second formula.

If both variables cross, the two single residues count the intersection
where \(z=x\) and \(w=y\) once too many.  The product of the residue signs
from \(1/(x-z)\) and \(1/(y-w)\), together with the overall factor
\(-\ii\) in \(\chi\), gives the contact term
\(\ii\mathcal L_g(x,y)\).  Subtracting this double residue gives the final
\(-\ii\mathcal L_g(x,y)\) contribution in
\eqref{eq:planar-n4-dhm-double-residue-continuation}.  The planar \(N=4\)
companion calculation check verifies these signs on finite Laurent models of
the DHM kernel, including the failure that occurs if the contact term is
omitted.
\end{proof}

\paragraph{DHM pole lattice and admissible local continuations.}
Let
\[
  \Delta(z,w)=z+z^{-1}-w-w^{-1}.
\]
For fixed \(g>0\), the Gamma-kernel logarithm \(\mathcal L_g(z,w)\) can
develop singularities only on the divisors
\begin{equation}
  \Delta(z,w)=\frac{\ii n}{g}
  \quad\hbox{or}\quad
  \Delta(z,w)=-\frac{\ii n}{g},
  \qquad n\in\mathbb Z_{\ge1}.
  \label{eq:planar-n4-dhm-pole-lattice}
\end{equation}
Consequently the one-residue correction \(\Psi_x^{\epsilon_y}(x,y)\) is a
single-valued holomorphic function on any connected region in which
\[
  x+x^{-1}-w-w^{-1}\notin \frac{\ii}{g}
  \bigl(\mathbb Z\setminus\{0\}\bigr)
\]
for every \(w\) on the chosen \(y\)-contour, and similarly for
\(\Psi_y^{\epsilon_x}\).  The contact term
\(\mathcal C_g(x,y)=\ii\mathcal L_g(x,y)\) is holomorphic precisely away
from the endpoint divisors
\[
  x+x^{-1}-y-y^{-1}\in
  \frac{\ii}{g}\bigl(\mathbb Z\setminus\{0\}\bigr).
\]
If a crossing path for \(x_1^\pm\) satisfies these avoidance conditions for
each of the four pairs \((x_1^\alpha,x_2^\beta)\),
\(\alpha,\beta\in\{+,-\}\), except for the Cauchy-pole crossings already
accounted for in
Proposition~\ref{prop:planar-n4-dhm-local-residue-continuation}, then
the local BES-continuation formula below applies with no additional
Gamma-pole residue terms.

The Gamma function has poles at \(0,-1,-2,\ldots\).  The numerator in
\(\mathcal L_g\) is singular only when
\[
  1+\ii g\Delta(z,w)=1-n,\qquad n\in\mathbb Z_{\ge1},
\]
which is equivalent to \(\Delta(z,w)=\ii n/g\).  The denominator is singular
only when
\[
  1-\ii g\Delta(z,w)=1-n,\qquad n\in\mathbb Z_{\ge1},
\]
which is equivalent to \(\Delta(z,w)=-\ii n/g\).  This proves
\eqref{eq:planar-n4-dhm-pole-lattice}.  The displayed conditions for
\(\Psi_x\), \(\Psi_y\), and \(\mathcal C_g\) are the same divisors evaluated
after setting \(z=x\), \(w=y\), or one variable on the remaining contour.
Away from those divisors the integrands are holomorphic in the external
variables and the contour deformations are ordinary Cauchy deformations.
Thus a path that crosses only the specified Cauchy poles has exactly the
residue contributions listed above, with no further Gamma-kernel terms.

\paragraph{Local crossing of the BES phase.}
Assume \(x_1^\pm\) and \(x_2^\pm\) lie in the physical DHM chamber
\(|x_1^\pm|>1\), \(|x_2^\pm|>1\), and continue the first magnon along the
crossing path of Definition~\ref{def:planar-n4-crossing-path}, so that
\(\bar x_1^\pm=1/x_1^\pm\) lie in the inside radial chamber.  Suppose the
path avoids the Gamma-function poles of the DHM kernel.  Define
\begin{equation}
\begin{aligned}
  \theta_{\rm raw}^{-,+}(\bar x_1^\pm,x_2^\pm)
  &=
  \chi_{\rm raw}^{-,+}(\bar x_1^+,x_2^+)
  -\chi_{\rm raw}^{-,+}(\bar x_1^+,x_2^-)
  -\chi_{\rm raw}^{-,+}(\bar x_1^-,x_2^+)
  +\chi_{\rm raw}^{-,+}(\bar x_1^-,x_2^-),\\
  \Delta_{\bar1}\theta
  &=
  -\Psi_x^{+}(\bar x_1^+,x_2^+)
  +\Psi_x^{+}(\bar x_1^+,x_2^-)
  +\Psi_x^{+}(\bar x_1^-,x_2^+)
  -\Psi_x^{+}(\bar x_1^-,x_2^-).
\end{aligned}
  \label{eq:planar-n4-crossed-bes-theta-local-residue}
\end{equation}
Then the local DHM continuation of the BES phase for the crossed first
particle is
\begin{equation}
  \theta_{\bar1 2}^{\rm loc}
  =
  \theta_{\rm raw}^{-,+}(\bar x_1^\pm,x_2^\pm)
  +
  \Delta_{\bar1}\theta .
  \label{eq:planar-n4-crossed-bes-theta-local}
\end{equation}

The BES phase is the alternating four-kernel combination
\[
  \chi(x_1^+,x_2^+)-\chi(x_1^+,x_2^-)
  -\chi(x_1^-,x_2^+)+\chi(x_1^-,x_2^-).
\]
Along the stated crossing path, only the first argument of each kernel crosses
from the outside chamber to the inside chamber; the second argument stays
outside.  The local DHM continuation formula for one kernel therefore adds
the correction \(-\Psi_x^+\) to each crossed kernel.  Multiplying these four
single-kernel corrections by the signs \(+,-,-,+\) in the BES combination
gives exactly the four terms defining \(\Delta_{\bar1}\theta\).  The raw
continued kernels give \(\theta_{\rm raw}^{-,+}\), so their sum is
\eqref{eq:planar-n4-crossed-bes-theta-local}.

This local contour computation fixes how the DHM
residue terms enter the four-kernel BES phase in the stringbook convention,
but it does not by itself prove the scalar crossing equation
\eqref{eq:planar-n4-crossing-scalar-factor}.  That equation also requires the
global branch of the scalar factor, the homotopy class of the crossing path,
and the absence or prescribed treatment of additional crossed singularities.

Inside this domain the weak-coupling coefficients can be extracted without
any sheet continuation.  Expand
\begin{equation}
  \chi(x,y)
  =
  \sum_{r=2}^{\infty}
  \sum_{\substack{s>r\\ r+s\ {\rm odd}}}
  \frac{c_{r,s}(g)}{(r-1)(s-1)}
  \left(
    \frac{1}{x^{r-1}y^{s-1}}
    -
    \frac{1}{x^{s-1}y^{r-1}}
  \right).
  \label{eq:planar-n4-chi-coefficient-expansion}
\end{equation}
The charge-expansion convention above uses the same \(c_{r,s}\); the two
forms agree because \(q_r=\ii A_r/(r-1)\), where
\(A_r=(x^+)^{1-r}-(x^-)^{1-r}\).

\begin{proposition}[Weak coefficients from the DHM contour]
\label{prop:planar-n4-dhm-weak-coefficients}
For \(N=2m+1\geq3\), let \(R^{(N)}_{r,s}\) be the Laurent coefficient
\[
  R^{(N)}_{r,s}
  =
  [z^{-(r-1)}w^{-(s-1)}]
  \bigl(z+z^{-1}-w-w^{-1}\bigr)^N .
\]
Then the contribution of order \(g^N\) to the coefficient \(c_{r,s}(g)\) in
\eqref{eq:planar-n4-chi-coefficient-expansion} is
\begin{equation}
  c_{r,s}(g)\big|_{g^N}
  =
  \frac{2(-1)^{m+1}}{N}\,
  (r-1)(s-1)R^{(N)}_{r,s}\,
  \zeta(N)g^N .
  \label{eq:planar-n4-dhm-residue-coefficient}
\end{equation}
Equivalently, writing
\[
  c_{r,s}(g)\big|_{g^N}=C^{(N)}_{r,s}\zeta(N)g^N,
\]
one has the closed coefficient
\begin{equation}
  C^{(N)}_{r,s}
  =
  \frac{
    2(-1)^{(N-1)/2+s}(r-1)(s-1)N!(N-1)!
  }{
    \Gamma\!\left(\frac{N-r-s+4}{2}\right)
    \Gamma\!\left(\frac{N+r-s+2}{2}\right)
    \Gamma\!\left(\frac{N-r+s+2}{2}\right)
    \Gamma\!\left(\frac{N+r+s}{2}\right)
  },
  \label{eq:planar-n4-dhm-closed-weak-coefficient}
\end{equation}
when the four Gamma-function arguments are positive integers; otherwise this
order gives zero.  In particular, the coefficient vanishes unless the
Laurent powers can be supplied by the \(N\) factors in
\((z+z^{-1}-w-w^{-1})^N\).
In particular,
\[
  c_{2,3}(g)=4\zeta(3)g^3-40\zeta(5)g^5+O(g^7).
\]
The first dressing contribution to a weak-coupling anomalous dimension is
nevertheless order \(g^6\), since \(q_2=O(g)\) and \(q_3=O(g^2)\) on the
physical weak-coupling branch.
\end{proposition}

\begin{proof}
For \(|x|,|y|>1\),
\[
  \frac{1}{x-z}\frac{1}{y-w}
  =
  \sum_{a,b\geq0}\frac{z^aw^b}{x^{a+1}y^{b+1}} .
\]
The logarithm in \eqref{eq:planar-n4-bes-chi-integral} has the Taylor
expansion at small \(g\)
\[
  \log\frac{\Gamma(1+\eta)}{\Gamma(1-\eta)}
  =
  -2\gamma\eta
  -
  2\sum_{m=1}^{\infty}
  \frac{\zeta(2m+1)}{2m+1}\eta^{2m+1},
  \qquad
  \eta=\ii g(z+z^{-1}-w-w^{-1}).
\]
The linear Euler-constant term is symmetric data that drops out of the
antisymmetric coefficient family \(s>r\).  For \(N=2m+1\geq3\), multiplying by
the outside factor \(-\ii\) gives
\[
  \frac{2(-1)^{m+1}\zeta(N)}{N}
  g^N
  \bigl(z+z^{-1}-w-w^{-1}\bigr)^N .
\]
The two contour integrals pick the coefficient of \(z^{-1}w^{-1}\).  Thus the
coefficient of \(x^{1-r}y^{1-s}\) is the displayed constant times
\(R^{(N)}_{r,s}\).  Comparing with
\eqref{eq:planar-n4-chi-coefficient-expansion} proves
\eqref{eq:planar-n4-dhm-residue-coefficient}.

It remains to evaluate the Laurent coefficient.  Put
\(a=r-1\) and \(b=s-1\).  Expanding first in the number \(\ell\) of
\((-w-w^{-1})\) factors gives the finite convolution
\begin{equation}
  R^{(N)}_{r,s}
  =
  \sum_{\ell=0}^{N}
  (-1)^\ell
  \binom{N}{\ell}
  \binom{N-\ell}{(N-\ell+a)/2}
  \binom{\ell}{(\ell+b)/2},
  \label{eq:planar-n4-dhm-residue-binomial-sum}
\end{equation}
where a binomial coefficient is declared zero unless its lower entry is an
integer between \(0\) and the upper entry.  This convention records the parity
and support restrictions.

For the closed evaluation, compare Fourier coefficients by writing
\(z=\ee^{\ii\theta}\), \(w=\ee^{\ii\varphi}\),
\(A=(\theta+\varphi)/2\), and \(B=(\theta-\varphi)/2\).  Then
\[
  z+z^{-1}-w-w^{-1}
  =
  -4\sin A\,\sin B,
  \qquad
  z^{-a}w^{-b}=\ee^{-\ii(a+b)A}\ee^{-\ii(a-b)B}.
\]
The elementary expansion
\[
  \sin^N t
  =
  (2\ii)^{-N}
  \sum_{j=0}^{N}
  (-1)^j\binom{N}{j}\ee^{\ii(N-2j)t}
\]
therefore gives
\[
  R^{(N)}_{r,s}
  =
  (-1)^{N+a}
  \binom{N}{(N+a+b)/2}
  \binom{N}{(N+a-b)/2},
\]
with the same convention that the binomial coefficient is zero unless its
lower entry is an integer between \(0\) and \(N\).  Since \(N\) is odd and the
nonzero cases have \(a+b\) odd, the sign is
\((-1)^{N+a}=(-1)^b=(-1)^{s-1}\).  Rewriting the two binomial coefficients as
Gamma functions gives
\[
  R^{(N)}_{r,s}
  =
  (-1)^{s-1}
  \frac{(N!)^2}{
    \Gamma\!\left(\frac{N-r-s+4}{2}\right)
    \Gamma\!\left(\frac{N+r-s+2}{2}\right)
    \Gamma\!\left(\frac{N-r+s+2}{2}\right)
    \Gamma\!\left(\frac{N+r+s}{2}\right)
  } .
\]
The lower-entry bounds in the binomial formula are exactly the positivity
conditions on the four Gamma-function arguments; outside those bounds the
coefficient is zero.  Substitution into
\eqref{eq:planar-n4-dhm-residue-coefficient} gives
\eqref{eq:planar-n4-dhm-closed-weak-coefficient}.  For \(r=2,s=3\),
\[
  R^{(3)}_{2,3}=3,\qquad R^{(5)}_{2,3}=50,
\]
which gives \(4\zeta(3)g^3-40\zeta(5)g^5\).
\end{proof}

A saddle expansion of the same contour representation gives the AFS and
Hernandez--Lopez strong-coupling terms.

\paragraph{Crossing scalar convention algebra.}
Assume \(x_1^\pm,x_2^\pm\) are non-zero and avoid the displayed
denominators.  In the stringbook scalar-factor orientation used in this
chapter set
\[
  \Xi_{12}
  =
  \frac{x_2^+}{x_2^-}
  \frac{x_2^+ -x_1^+}{x_2^+ -x_1^-}
  \frac{x_2^- -1/x_1^+}{x_2^- -1/x_1^-}.
\]
The common reciprocal orientation is
\[
  \Xi_{12}^{\rm rec}
  =
  \frac{x_2^-}{x_2^+}
  \frac{x_2^+ -x_1^-}{x_2^+ -x_1^+}
  \frac{x_2^- -1/x_1^-}{x_2^- -1/x_1^+},
\]
and satisfies
\[
  \Xi_{12}^{\rm rec}=\Xi_{12}^{-1}.
\]
This is the product of the three displayed inverse factors:
\[
\begin{aligned}
  \Xi_{12}\Xi_{12}^{\rm rec}
  &=
  \frac{x_2^+}{x_2^-}\frac{x_2^-}{x_2^+}
  \frac{x_2^+ -x_1^+}{x_2^+ -x_1^-}
  \frac{x_2^+ -x_1^-}{x_2^+ -x_1^+}
  \frac{x_2^- -1/x_1^+}{x_2^- -1/x_1^-}
  \frac{x_2^- -1/x_1^-}{x_2^- -1/x_1^+}
  =1 .
\end{aligned}
\]
Moreover the algebraic substitution \(x_1^\pm\mapsto1/x_1^\pm\) does not
leave this multiplier invariant as a rational function.  At
\[
  x_1^+=2,\qquad x_1^-=3,\qquad x_2^+=5,\qquad x_2^-=7
\]
one obtains \(\Xi_{12}=117/112\), whereas the same formula with
\(x_1^\pm\) replaced by \(1/x_1^\pm\) gives \(675/784\).  Hence equality can
hold only on a proper algebraic sublocus, not as a convention-independent
identity.

\paragraph{Matrix-channel origin of the scalar crossing multiplier.}
Let \(\bar x_1^\pm=1/x_1^\pm\) denote the crossed-sheet continuation of the
first magnon and assume all displayed denominators are non-zero.  Form the
crossed amplitude \(L_{\bar1 2}\) from
\eqref{eq:planar-n4-su2c-ten-amplitudes} by replacing
\((x_1^+,x_1^-)\) with \((\bar x_1^+,\bar x_1^-)\), and form \(G_{21}\) by
exchanging the two magnons and using the analytic-unitarity convention
\(\mathcal S_{21}=\mathcal S_{12}^{-1}\).  Then
\begin{equation}
  \mathcal R_{\rm cr}
  :=
  \frac{L_{\bar1 2}}{G_{21}}
  =
  \mathcal S_{\bar1 2}\mathcal S_{12}
  \frac{(x_2^- -1/x_1^-)\,x_1^+(x_1^- -x_2^+)}
       {(x_2^-x_1^+-1)(x_1^+ -x_2^+)} .
  \label{eq:planar-n4-crossing-channel-ratio}
\end{equation}
Substitute the stringbook scalar split for
\(\mathcal S_{12}\) and \(\mathcal S_{\bar1 2}\).  In the chamber
\[
  x_2^->x_1^+>1,\qquad x_2^+>x_1^->1,
\]
where the square roots in the scalar split are taken continuously from
positive real values, the finite channel condition
\(\mathcal R_{\rm cr}=1\) is equivalent to
\begin{equation}
  \sigma(\bar x_1^\pm,x_2^\pm)\sigma(x_1^\pm,x_2^\pm)=\Xi_{12}.
  \label{eq:planar-n4-crossing-sigma-from-channel}
\end{equation}
The formula for \(L_{\bar1 2}\) is
\[
  L_{\bar1 2}
  =
  \mathcal S_{\bar1 2}
  \frac{x_2^- -1/x_1^-}{x_2^- -1/x_1^+},
\]
whereas the exchanged \(G\)-amplitude is
\[
  G_{21}
  =
  \mathcal S_{21}
  \frac{x_1^+ -x_2^+}{x_1^- -x_2^+}.
\]
Using \(\mathcal S_{21}=\mathcal S_{12}^{-1}\) and rewriting
\((x_2^- -1/x_1^+)=(x_2^-x_1^+-1)/x_1^+\) gives
\eqref{eq:planar-n4-crossing-channel-ratio}.

Now write
\[
  \mathcal S_{12}=R_{12}^{1/2}\sigma_{12},
  \qquad
  \mathcal S_{\bar1 2}=R_{\bar1 2}^{1/2}\sigma_{\bar1 2},
\]
with \(R\) the rational factor displayed in the scalar split.  Squaring the
condition \(\mathcal R_{\rm cr}=1\) removes the square-root branch and gives
a rational identity:
\[
  R_{12}R_{\bar1 2}\Xi_{12}^{2}
  \left[
    \frac{(x_2^- -1/x_1^-)\,x_1^+(x_1^- -x_2^+)}
         {(x_2^-x_1^+-1)(x_1^+ -x_2^+)}
  \right]^2
  =1 .
\]
After cancellation this identity is exactly the product of the factors in
\(\Xi_{12}\).  The stated real chamber fixes the sign of the square root, so
the unsquared channel equation is
\eqref{eq:planar-n4-crossing-sigma-from-channel}.  This is the finite
matrix-amplitude calculation behind the scalar multiplier; it does not
construct the analytic crossed scalar branch.

\begin{proposition}[Scalar crossing monodromy cocycle]
\label{prop:planar-n4-crossing-scalar-monodromy-cocycle}
Let \(\mathsf C_1\) denote a chosen crossing continuation of the first
magnon along \(\gamma_{\rm cr}\) from
Definition~\ref{def:planar-n4-crossing-path}.  Fix a scalar branch
\(\sigma^{[\gamma]}_{12}\) on a sheet reached by a path \(\gamma\) from the
physical chamber, and write
\[
  \sigma^{[\gamma\mathsf C_1]}_{12}
  :=
  \sigma^{[\gamma\mathsf C_1]}(x_1^\pm,x_2^\pm)
\]
for the same external coordinates evaluated after analytic continuation
along the concatenated path.  Assume that the crossed path avoids all poles
and zeros in the denominators below and that Janik crossing holds on each
crossed segment in the stringbook orientation,
\[
  \sigma^{[\gamma\mathsf C_1]}_{12}\,
  \sigma^{[\gamma]}_{12}
  =\Xi_{12}(x_1^\pm,x_2^\pm).
\]
Then the second crossing is not an involution on the scalar branch.  It acts
by
\begin{equation}
  \frac{\sigma^{[\gamma\mathsf C_1\mathsf C_1]}_{12}}
       {\sigma^{[\gamma]}_{12}}
  =
  \frac{\Xi_{12}(1/x_1^\pm,x_2^\pm)}
       {\Xi_{12}(x_1^\pm,x_2^\pm)}
  =:\mathcal M_{12}^{(2)},
  \label{eq:planar-n4-crossing-double-monodromy}
\end{equation}
where explicitly
\begin{equation}
  \mathcal M_{12}^{(2)}
  =
  \frac{x_2^+ -1/x_1^+}{x_2^+ -1/x_1^-}
  \frac{x_2^+ -x_1^-}{x_2^+ -x_1^+}
  \frac{x_2^- -x_1^+}{x_2^- -x_1^-}
  \frac{x_2^- -1/x_1^-}{x_2^- -1/x_1^+}.
  \label{eq:planar-n4-crossing-double-monodromy-expanded}
\end{equation}
In the reciprocal scalar convention
\(\sigma^{\rm rec}_{12}:=\sigma_{12}^{-1}\), the one-crossing multiplier
and the double-crossing multiplier are inverted.
\end{proposition}

\begin{proof}
The first crossing gives
\[
  \sigma^{[\gamma\mathsf C_1]}_{12}
  =
  \frac{\Xi_{12}(x_1^\pm,x_2^\pm)}
       {\sigma^{[\gamma]}_{12}} .
\]
After that continuation the projected Zhukovsky variables of the first
magnon are \(1/x_1^\pm\).  Applying the same crossing equation once more
therefore gives
\[
  \sigma^{[\gamma\mathsf C_1\mathsf C_1]}_{12}
  =
  \frac{\Xi_{12}(1/x_1^\pm,x_2^\pm)}
       {\sigma^{[\gamma\mathsf C_1]}_{12}}
  =
  \frac{\Xi_{12}(1/x_1^\pm,x_2^\pm)}
       {\Xi_{12}(x_1^\pm,x_2^\pm)}
  \sigma^{[\gamma]}_{12}.
\]
Substituting the definition of \(\Xi_{12}\) cancels the common
\(x_2^+/x_2^-\) factor and yields
\eqref{eq:planar-n4-crossing-double-monodromy-expanded}.  If the reciprocal
scalar factor is used, every displayed crossing equation is inverted because
\(\sigma^{\rm rec}_{12}=1/\sigma_{12}\); the same two-line calculation then
gives \((\mathcal M_{12}^{(2)})^{-1}\).  The argument proves the rational
branch displacement forced by the chosen crossing law.  It does not classify
all possible homotopy classes of scalar branches or prove that the BES
branch is the unique minimal solution.
\end{proof}

\begin{proposition}[CDD ambiguity of the scalar crossing problem]
\label{prop:planar-n4-crossing-cdd-ambiguity}
Let \(\mathcal U\) be a regular domain of scalar branches closed under
particle exchange and under the chosen first-particle crossing segments
\(\mathsf C_1\), with no zero or pole of the functions below on the crossed
segments.  Suppose \(\sigma_{12}\) and \(\sigma'_{12}\) are two non-zero
meromorphic scalar branches on \(\mathcal U\) satisfying the same Janik
crossing equation
\[
  \sigma^{[\gamma\mathsf C_1]}_{12}\,
  \sigma^{[\gamma]}_{12}
  =
  \sigma'^{[\gamma\mathsf C_1]}_{12}\,
  \sigma'^{[\gamma]}_{12}
  =
  \Xi_{12}
\]
and the same scalar unitarity equation
\[
  \sigma^{[\gamma P]}_{21}\sigma^{[\gamma]}_{12}
  =
  \sigma'^{[\gamma P]}_{21}\sigma'^{[\gamma]}_{12}
  =1,
\]
where \(P\) exchanges the two particles.  Set
\[
  \mathcal F^{[\gamma]}_{12}
  :=
  \frac{\sigma'^{[\gamma]}_{12}}{\sigma^{[\gamma]}_{12}} .
\]
Then \(\mathcal F\) is a CDD factor in this convention:
\begin{equation}
  \mathcal F^{[\gamma\mathsf C_1]}_{12}
  \mathcal F^{[\gamma]}_{12}=1,\qquad
  \mathcal F^{[\gamma P]}_{21}
  \mathcal F^{[\gamma]}_{12}=1,\qquad
  \mathcal F^{[\gamma\mathsf C_1\mathsf C_1]}_{12}
  =
  \mathcal F^{[\gamma]}_{12}.
  \label{eq:planar-n4-cdd-factor-equations}
\end{equation}
Conversely, multiplying any scalar solution by a non-zero meromorphic
\(\mathcal F\) satisfying the first two equations in
\eqref{eq:planar-n4-cdd-factor-equations} gives another scalar solution with
the same crossing multiplier and scalar unitarity.  If \(\mathcal F\) has a
zero of order \(m\) at a regular point of a branch, then it has a pole of
order \(m\) at the corresponding crossed point and at the exchanged point.
Thus a CDD-free BES branch is a minimality and analyticity choice, not a
formal consequence of the finite matrix crossing multiplier.
\end{proposition}

\begin{proof}
Dividing the crossing equation for \(\sigma'\) by the crossing equation for
\(\sigma\) gives
\[
  \mathcal F^{[\gamma\mathsf C_1]}_{12}
  \mathcal F^{[\gamma]}_{12}=1.
\]
Dividing scalar unitarity for \(\sigma'\) by scalar unitarity for
\(\sigma\) gives
\[
  \mathcal F^{[\gamma P]}_{21}
  \mathcal F^{[\gamma]}_{12}=1.
\]
Applying the first identity on the already crossed branch gives
\[
  \mathcal F^{[\gamma\mathsf C_1\mathsf C_1]}_{12}
  \mathcal F^{[\gamma\mathsf C_1]}_{12}=1.
\]
Together with \(\mathcal F^{[\gamma\mathsf C_1]}_{12}
=(\mathcal F^{[\gamma]}_{12})^{-1}\), this proves the double-crossing
identity for the CDD factor.  Conversely, if
\(\sigma'_{12}:=\mathcal F_{12}\sigma_{12}\), then the first CDD equation
cancels the extra factor in crossing and the second cancels the extra factor
in scalar unitarity.

For the divisor statement, work in a coordinate chart where crossing and
exchange are biholomorphic.  The identity
\(\mathcal F^{[\gamma\mathsf C_1]}=1/\mathcal F^{[\gamma]}\) changes an
order-\(m\) zero into an order-\(m\) pole on the crossed branch, and the
exchange identity does the same on the exchanged branch.  In a simply
connected pole-free subdomain one may write
\(\mathcal F=\exp \ell\), and the equations become
\[
  \ell^{[\gamma\mathsf C_1]}_{12}=-\ell^{[\gamma]}_{12},
  \qquad
  \ell^{[\gamma P]}_{21}=-\ell^{[\gamma]}_{12}.
\]
Hence setting \(\ell=0\), or equivalently \(\mathcal F=1\), is an additional
minimal-branch normalization.
\end{proof}

Crossing relates \(\sigma(x_1^\pm,x_2^\pm)\) to the phase obtained by
analytically continuing the first magnon through the crossed sheet
\(x_1^\pm\mapsto1/x_1^\pm\).  It supplies an additional analyticity condition
on the scalar factor beyond factorized Yang--Baxter compatibility.
The preceding propositions fix the rational convention carried by the
chosen scalar split, its finite crossed-amplitude origin, and the
double-crossing sheet displacement forced by that convention, while
\eqref{eq:planar-n4-cdd-factor-equations} identifies the remaining
CDD freedom.  They do not derive the global monodromy condition from finite
algebra.  Homotopic crossing paths inside a singularity-free chamber give
the same rational cocycle, but paths separated by poles, zeros, or extra CDD
singularities can land on different scalar branches.  In the
normalization where the rational matrix part is separated as
above, Janik crossing has the convention-fixed form
\begin{equation}
  \sigma(\bar x_1^\pm,x_2^\pm)\sigma(x_1^\pm,x_2^\pm)
  =
  \frac{x_2^+}{x_2^-}
  \frac{x_2^+ -x_1^+}{x_2^+ -x_1^-}
  \frac{x_2^- -1/x_1^+}{x_2^- -1/x_1^-} ,
  \label{eq:planar-n4-crossing-scalar-factor}
\end{equation}
where \(\bar x_1^\pm\) denotes the crossed-sheet continuation of the first
magnon on the same rapidity surface.  The equation is best read as a monodromy
statement for the scalar factor on the
Zhukovsky rapidity surface.  The BES solution selects the minimal solution
compatible with \eqref{eq:planar-n4-crossing-scalar-factor}, unitarity
\(\sigma_{12}\sigma_{21}=1\), and the prescribed weak/strong asymptotics.

\begin{definition}[Dressing-phase data and status]
\label{def:planar-n4-dressing-phase-data}
The scalar dressing phase used in the asymptotic equations is specified by
the antisymmetric charge expansion above, the AFS leading strong-coupling
term \(c_{r,r+1}(g)=g+O(g^0)\), the Hernandez--Lopez semiclassical
correction, Janik crossing on the Zhukovsky surface, and the BES integral
representation selecting the minimal crossing-symmetric solution.  Its weak
series begins with
\[
  c_{2,3}(g)=4\zeta(3)g^3+O(g^5),
\]
and, because the weak charges contribute \(q_2q_3-q_3q_2=O(g^3)\), the first
dressing contribution to Konishi enters at four loops.  This is an exact
datum of the planar-integrability framework and a highly tested claim about
gauge theory, not a theorem here derived from a constructive definition of
four-dimensional \(\mathcal N=4\) SYM.
\end{definition}

\begin{quotedtheorem}[Asymptotic planar magnon S-matrix]
\label{qthm:planar-n4-asymptotic-smatrix-status}
Assume planar integrability of the long \(\mathcal N=4\) SYM spin chain, the
centrally extended residual symmetry, physical unitarity, Yang--Baxter
factorization, crossing, and the minimal analyticity condition used in the
BES construction.  Then the asymptotic two-body magnon S-matrix is the
\(\mathfrak{su}(2|2)^2\)-invariant matrix part multiplied by the scalar
dressing factor whose phase has the charge expansion above with the BES
coefficients \(c_{r,s}(g)\).
\end{quotedtheorem}

This is stated as a quoted theorem because the full claim is not a theorem of
four-dimensional QFT derived from a nonperturbative definition of the gauge
theory.  It is an integrability theorem inside a framework whose physical
identification with the planar gauge-theory dilatation operator has extensive
checks and strong structural constraints.

\section{Asymptotic Bethe Equations}

\begin{assumption}[Asymptotic Bethe--Yang regime]
\label{ass:planar-n4-aba-regime}
The all-loop asymptotic Bethe ansatz is used as a long-chain
Bethe--Yang quantization rule for planar single traces.  More precisely, for
a sequence of operators of length \(L\), with either fixed magnon number or a
controlled large-\(L\) distribution of Bethe roots, one assumes:
\[
  L\to\infty,\qquad
  \text{virtual propagation around the trace is suppressed,}
\]
and the many-magnon quantization is obtained by factorizing scattering into
the two-body S-matrix of the preceding section.  This intended domain includes
high-density configurations of roots in the \(L\to\infty\) limit.  The
additional assumption is not thermodynamic Bethe ansatz; it is the
asymptotic finite-volume quantization of the physical spin-chain problem.
At finite \(L\), or at a loop order where the long-range dilatation operator
can wrap around the trace, the omitted mirror-winding effects are no longer
parametrically absent.
\end{assumption}

The word ``asymptotic'' in this assumption refers to the spatial length of the
spin chain, not to a sparse-gas approximation.  The following elementary
counting calculation is the precise large-\(L\) bridge used whenever an ABA
solution is replaced by a macroscopic root distribution.

\paragraph{Finite-density ABA counting equation.}
Consider a diagonal component of a long-chain Bethe--Yang problem with rapidity
variables \(u_j\), \(j=1,\ldots,M_L\), and logarithmic equations
\begin{equation}
  Lp(u_j)+\sum_{\substack{k=1\\k\neq j}}^{M_L}\Theta(u_j,u_k)
  =
  2\pi I_j,
  \qquad I_j\in\mathbb Z .
  \label{eq:planar-n4-aba-log-counting-form}
\end{equation}
Assume a branch of the two-body phase \(\Theta\) has been chosen on the
rapidity interval under consideration; possible diagonal jumps have either
been removed by this branch choice or are treated as explicit source terms.
Assume also that \(M_L/L\) has a finite limit and that the empirical measures
\[
  \nu_L=\frac1L\sum_{j=1}^{M_L}\delta_{u_j}
\]
converge weakly to \(\rho(v)\dd v\) on compact subintervals where
\(p\) and \(\Theta\) are \(C^1\).  Define
\[
  Z_L(u)=p(u)+\frac1L\sum_{j=1}^{M_L}\Theta(u,u_j),
  \qquad
  Z(u)=p(u)+\int\Theta(u,v)\rho(v)\,\dd v .
\]
Then \(Z_L\to Z\) locally uniformly after differentiating under the stated
regularity hypotheses, and the density of available Bethe levels is
\begin{equation}
  \rho(u)+\rho_{\rm h}(u)
  =
  \frac{1}{2\pi}Z'(u)
  =
  \frac{p'(u)}{2\pi}
  +
  \int K(u,v)\rho(v)\,\dd v,
  \qquad
  K(u,v)=\frac{1}{2\pi}\partial_u\Theta(u,v).
  \label{eq:planar-n4-aba-finite-density-counting-equation}
\end{equation}
Here \(\rho_{\rm h}\) is the macroscopic density of unoccupied integer levels
in the chosen branch.  In a completely filled interval \(\rho_{\rm h}=0\).

Weak convergence gives
\[
  \frac1L\sum_{j=1}^{M_L}\Theta(u,u_j)
  \longrightarrow
  \int\Theta(u,v)\rho(v)\,\dd v
\]
uniformly on compact subsets on which the phase is continuous and the root
cloud stays away from unaccounted singularities.  The same statement applied
to \(\partial_u\Theta\) gives the derivative formula for \(Z'(u)\).
Equation~\eqref{eq:planar-n4-aba-log-counting-form} says that the integer
level attached to rapidity \(u\) is, to leading order, \(LZ(u)/(2\pi)\).  Hence
the number of available integer levels in a small interval \([u,u+\dd u]\) is
\[
  \frac{L}{2\pi}Z'(u)\,\dd u+o(L\,\dd u).
\]
Dividing by \(L\,\dd u\) gives the density of occupied plus unoccupied
available levels, namely \(\rho+\rho_{\rm h}\).  This proves
\eqref{eq:planar-n4-aba-finite-density-counting-equation}.

This is a large-\(L\) continuum limit of the same physical ABA equations.  It
is not mirror TBA: no mirror energy, entropy functional, pseudoenergy, or
thermal variational problem has been introduced.  The additional dynamical
assumption is exactly the one stated above, namely that the long-chain
two-body magnon S-matrix continues to quantize the physical chain when
\(M_L/L=O(1)\), as long as mirror winding remains asymptotically suppressed.

For the rank-one equation in the stringbook \(SL(2)\)-vacuum orientation, the
all-loop asymptotic Bethe equations can be written
\begin{equation}
  \left(\frac{x_j^+}{x_j^-}\right)^L
  =
  \prod_{\substack{k=1\\ k\neq j}}^M
  \frac{x_j^- -x_k^+}{x_j^+ -x_k^-}
  \frac{1-1/(x_j^+x_k^-)}{1-1/(x_j^-x_k^+)}
  \sigma(x_j^\pm,x_k^\pm)^2 ,
  \label{eq:planar-n4-rank-one-stringbook-aba}
\end{equation}
with cyclicity
\[
  \prod_{j=1}^M\frac{x_j^+}{x_j^-}=1.
\]
The scaling dimension is
\[
  \Delta
  =
  L+\sum_{j=1}^M\left(
    \sqrt{1+16g^2\sin^2\frac{p_j}{2}}-1
  \right),
\]
up to wrapping corrections.  The full
\(\mathfrak{psu}(2,2|4)\) asymptotic Bethe equations require nested roots
for the residual symmetry algebra.  Their algebraic role is the same as in
Chapter~\ref{ch:integrable-nested-bethe-ansatz-matrix-bethe-yang}; their
physical meaning is different because these auxiliary roots organize
polarizations of magnons on a gauge-theory trace.

The first nesting step can be derived without the full
\(\mathfrak{psu}(2,2|4)\) bookkeeping.  Start with \(M\) level-I magnons
\(\phi_1(p_k)\).  A single level-II fermionic excitation with spectral
parameter \(y\) is a wave packet
\[
  \ket{\Psi_{\rm II}(y)}
  =
  \sum_{k=1}^M
  C_k(y;\{p_j\})\,
  \ket{\phi_1(p_1)\cdots\psi(p_k)\cdots\phi_1(p_M)} ,
\]
where the omitted flavour index of \(\psi\) is a spectator.  Write
\(x_i^\pm=x^\pm(p_i)\) and let \(\mathsf a_i=\mathsf a(p_i)\) be the
nonzero representation coefficient which converts a bosonic vacuum magnon
to the chosen fermionic excitation in the spin-chain basis.  Normalize the
two-body matrix part by the boson-boson amplitude \(A_{12}\).  In the
stringbook orientation of the \(\mathfrak{su}(2|2)_{\rm c}\) intertwiner,
the four mixed amplitudes needed for one fermionic defect obey
\begin{equation}
\begin{aligned}
  \frac{G_{12}}{A_{12}}
  &=
  \frac{x_2^+-x_1^+}{x_2^+-x_1^-},
  &
  \frac{K_{12}}{A_{12}}
  &=
  \frac{\mathsf a_2}{\mathsf a_1}
  \frac{x_1^+-x_1^-}{x_2^+-x_1^-},\\
  \frac{L_{12}}{A_{12}}
  &=
  \frac{x_2^--x_1^-}{x_2^+-x_1^-},
  &
  \frac{H_{12}}{A_{12}}
  &=
  \frac{\mathsf a_1}{\mathsf a_2}
  \frac{x_2^+-x_2^-}{x_2^+-x_1^-}.
\end{aligned}
  \label{eq:planar-n4-mixed-amplitude-ratios-for-nesting}
\end{equation}
These are ratios of matrix amplitudes, so the scalar dressing factor and
reciprocal scalar-factor conventions cancel.

\begin{proposition}[Single level-II nesting step]
\label{prop:planar-n4-single-level-ii-nesting-step}
Assume \(x_i^\pm\), \(\mathsf a_i\), and the denominators in
\eqref{eq:planar-n4-mixed-amplitude-ratios-for-nesting} are nonzero.  The
one-defect ansatz
\[
  C_k(y;\{p_j\})
  =
  f(y,p_k)\prod_{\ell<k}S^{{\rm II},{\rm I}}(y,p_\ell)
\]
is compatible with every adjacent transposition of level-I magnons if
\begin{equation}
  f(y,p)=\frac{\mathsf a(p)}{y-x^+(p)},
  \qquad
  S^{{\rm II},{\rm I}}(y,p)
  =
  \frac{y-x^-(p)}{y-x^+(p)} .
  \label{eq:planar-n4-single-level-ii-data}
\end{equation}
The normalization of \(f\) may be multiplied by a function of \(y\) alone;
this rescales the one-defect state and has no Bethe-equation content.
\end{proposition}

\begin{proof}
It is enough to check a two-site exchange, because the many-site statement is
then obtained by multiplying the same local move along the chain.  Let
\[
  f_i=f(y,p_i),
  \qquad
  S_i=S^{{\rm II},{\rm I}}(y,p_i).
\]
Before exchanging magnons \(1\) and \(2\), the defect can sit on the first
or second site with coefficients \(f_1\) and \(f_2S_1\).  After applying the
matrix S-matrix to this two-site block, equality with the ansatz in the
exchanged order gives the two coefficient equations
\begin{equation}
\begin{aligned}
  f_1L_{12}+f_2S_1H_{12}
  &=
  S_2f_1A_{12},\\
  f_1K_{12}+f_2S_1G_{12}
  &=
  f_2A_{12}.
\end{aligned}
  \label{eq:planar-n4-single-level-ii-local-equations}
\end{equation}
These equations are the algebraic content of moving a single auxiliary
fermion through two neighbouring physical magnons.

Substitute \eqref{eq:planar-n4-mixed-amplitude-ratios-for-nesting} and
\eqref{eq:planar-n4-single-level-ii-data}.  The second equation in
\eqref{eq:planar-n4-single-level-ii-local-equations} reduces, after
clearing the common factor
\(A_{12}\mathsf a_2/((y-x_1^+)(x_2^+-x_1^-))\), to
\[
  (x_1^+-x_1^-)(y-x_2^+)
  +(x_2^+-x_1^+)(y-x_1^-)
  =
  (x_2^+-x_1^-)(y-x_1^+).
\]
The first equation reduces in the same way to
\[
  (x_2^--x_1^-)(y-x_2^+)
  +(x_2^+-x_2^-)(y-x_1^-)
  =
  (x_2^+-x_1^-)(y-x_2^-).
\]
Both are polynomial identities in \(y\).  Therefore the local exchange is
consistent, and repeated adjacent exchanges give the displayed coefficient
formula for an arbitrary ordering of the level-I magnons.
\end{proof}

The square-root braiding factors of the string basis multiply
\eqref{eq:planar-n4-single-level-ii-data} by convention-dependent one-body
factors; they cancel from the local identities above after the corresponding
basis change is applied to the mixed amplitudes.

With two level-II fermions there is a genuine contact term: when the two
auxiliary fermions occupy the same level-I site, the local state can turn
into the boson \(\phi_2\) accompanied by the length-changing marker \(Z^+\).
Let
\[
  v(y)=y+y^{-1}.
\]
In the spin-chain basis, the local two-defect calculation fixes the
\(SU(2)\)-invariant level-II matrix
\begin{equation}
  (S^{{\rm II}})_{\alpha\beta}^{\delta\gamma}(y_1,y_2)
  =
  \delta_\alpha^\delta\delta_\beta^\gamma\,M(y_1,y_2)
  +
  \delta_\alpha^\gamma\delta_\beta^\delta\,N(y_1,y_2),
  \label{eq:planar-n4-level-ii-mn-definition}
\end{equation}
with
\begin{equation}
  M+N=-1,
  \qquad
  M-N=
  \frac{v(y_1)-v(y_2)-\ii/g}{v(y_1)-v(y_2)+\ii/g}.
  \label{eq:planar-n4-level-ii-mn-eigenvalues}
\end{equation}
Equivalently,
\[
  M(y_1,y_2)=
  -\,\frac{\ii/g}{v(y_1)-v(y_2)+\ii/g},
  \qquad
  N(y_1,y_2)=
  -\,\frac{v(y_1)-v(y_2)}{v(y_1)-v(y_2)+\ii/g}.
\]
The accompanying same-site coefficient is
\begin{equation}
  \mathsf f^{(2)}(y_1,y_2;p)
  =
  \frac{g}{\mathsf a(p)^2}
  \left(\frac{x^-(p)}{x^+(p)}-1\right)
  \frac{y_1y_2-x^+(p)x^-(p)}{y_1y_2}
  \frac{y_1-y_2}{v(y_1)-v(y_2)+\ii/g}.
  \label{eq:planar-n4-level-ii-contact-coefficient}
\end{equation}
The sign \(M+N=-1\) is not a scalar scattering phase to be inserted into
the level-III vacuum Bethe equation.  It is the ungraded matrix eigenvalue
on the state with two identical level-II fermions; after the fermionic
exchange sign is accounted for, the level-III vacuum scattering is trivial.
This is the local sign convention behind the sentence that the
\(\psi_1\)-vacuum has \(S^{{\rm II},{\rm II}}=1\).

\paragraph{Single level-III nesting step.}
\label{par:planar-n4-single-level-iii-nesting-step}
Assume \(g\neq0\), \(v(y_1)-v(y_2)+\ii/g\neq0\), and the denominators below
are nonzero.  In the graded convention just described, a single level-III
excitation with spectral parameter \(w\) has one-body coefficient
\[
  \gamma(w,y)=
  \frac{w+\ii/(2g)}{w-v(y)+\ii/(2g)}
\]
and level-III/level-II scattering factor
\begin{equation}
  S^{{\rm III},{\rm II}}(w,y)
  =
  \frac{w-v(y)-\ii/(2g)}
       {w-v(y)+\ii/(2g)}.
  \label{eq:planar-n4-level-iii-level-ii-factor}
\end{equation}

This is a two-site compatibility calculation after the level-II matrix,
contact-term convention, and graded exchange sign have been fixed.  Put
\(\eta=\ii/(2g)\), \(v_i=v(y_i)\), \(D=v_1-v_2\), and
\(a_i=w-v_i\).  Then \(a_2=a_1+D\), and
\eqref{eq:planar-n4-level-ii-mn-eigenvalues} is equivalent to
\[
  M=-\frac{2\eta}{D+2\eta},
  \qquad
  N=-\frac{D}{D+2\eta}.
\]
For an adjacent exchange of two level-II sites, the two independent
coefficients must obey
\begin{equation}
\begin{aligned}
  \gamma(w,y_1)M
  +\gamma(w,y_2)S^{{\rm III},{\rm II}}(w,y_1)N
  &=
  -\gamma(w,y_2),\\
  \gamma(w,y_1)N
  +\gamma(w,y_2)S^{{\rm III},{\rm II}}(w,y_1)M
  &=
  -\gamma(w,y_1)S^{{\rm III},{\rm II}}(w,y_2).
\end{aligned}
  \label{eq:planar-n4-level-iii-local-equations}
\end{equation}
The minus signs are precisely the fermionic exchange signs which convert the
ungraded \(M+N=-1\) vacuum eigenvalue into trivial graded vacuum scattering.

Substituting \(\gamma(w,y_i)=(w+\eta)/(a_i+\eta)\) and
\(S^{{\rm III},{\rm II}}(w,y_i)=(a_i-\eta)/(a_i+\eta)\), the first equation
in \eqref{eq:planar-n4-level-iii-local-equations} reduces to
\[
  -2\eta(a_2+\eta)-D(a_1-\eta)
  =
  -(D+2\eta)(a_1+\eta).
\]
The second reduces to
\[
  -D(a_2+\eta)-2\eta(a_1-\eta)
  =
  -(D+2\eta)(a_2-\eta).
\]
Both identities follow immediately from \(a_2=a_1+D\).  Therefore the
one-level-III ansatz is compatible with every adjacent level-II exchange.

A pair of level-III roots scatters by
\[
  S^{{\rm III},{\rm III}}(w_1,w_2)
  =
  \frac{w_1-w_2+\ii/g}{w_1-w_2-\ii/g}.
\]

\begin{proposition}[Single-copy nested Bethe--Yang equations]
\label{prop:planar-n4-single-copy-nested-bethe-yang}
In the spin-chain basis for one \(\mathfrak{su}(2|2)_{\rm c}\) copy, let
\(K^{\rm I}\), \(K^{\rm II}\), and \(K^{\rm III}\) be the numbers of
level-I momenta \(p_k\), level-II roots \(y_a\), and level-III roots \(w_b\).
Assume the asymptotic long-chain regime and the absence of singular
coincidences among the displayed factors.  The closed-chain periodicity
conditions are
\begin{equation}
\begin{aligned}
  \ee^{\ii p_kL}
  &=
  \prod_{\substack{j=1\\j\neq k}}^{K^{\rm I}}
  S^{{\rm I},{\rm I}}(p_j,p_k)
  \prod_{a=1}^{K^{\rm II}}
  S^{{\rm II},{\rm I}}(y_a,p_k),
  \qquad
  k=1,\ldots,K^{\rm I},\\
  1
  &=
  \prod_{j=1}^{K^{\rm I}}
  S^{{\rm II},{\rm I}}(y_a,p_j)^{-1}
  \prod_{b=1}^{K^{\rm III}}
  S^{{\rm III},{\rm II}}(w_b,y_a),
  \qquad
  a=1,\ldots,K^{\rm II},\\
  1
  &=
  \prod_{a=1}^{K^{\rm II}}
  S^{{\rm III},{\rm II}}(w_b,y_a)^{-1}
  \prod_{\substack{c=1\\c\neq b}}^{K^{\rm III}}
  S^{{\rm III},{\rm III}}(w_c,w_b),
  \qquad
  b=1,\ldots,K^{\rm III}.
\end{aligned}
  \label{eq:planar-n4-single-copy-nested-bethe-yang}
\end{equation}
For a state with \(N_1,N_2,N_3,N_4\) excitations of types
\(\phi_1,\phi_2,\psi_1,\psi_2\), the nesting numbers are
\[
  K^{\rm I}=N_1+N_2+N_3+N_4,\qquad
  K^{\rm II}=2N_2+N_3+N_4,\qquad
  K^{\rm III}=N_2+N_4.
\]
\end{proposition}

\begin{proof}
The first equation is obtained by transporting the \(k\)-th level-I magnon
once around the trace.  It crosses every other physical magnon and every
level-II defect, hence the two products in the first line.  The second
equation transports the \(a\)-th level-II defect around the level-II chain.
Crossing a physical magnon in the reverse direction contributes
\(S^{{\rm II},{\rm I}}(y_a,p_j)^{-1}\), while crossing a level-III defect
contributes \(S^{{\rm III},{\rm II}}(w_b,y_a)\).  The third equation
transports the \(b\)-th level-III defect around the level-III chain: each
level-II defect is crossed in the reverse direction, giving
\(S^{{\rm III},{\rm II}}(w_b,y_a)^{-1}\), and each other level-III defect is
crossed in the forward direction, giving
\(S^{{\rm III},{\rm III}}(w_c,w_b)\).  The three equations are therefore
the Bethe--Yang periodicity conditions for the three Zamolodchikov--Faddeev
chains, not thermodynamic Bethe ansatz.  The formulas for
\(K^{\rm I},K^{\rm II},K^{\rm III}\) follow from the nesting content of each
letter: \(\phi_1\) contributes only level I, \(\psi_1\) contributes one
level-II root, \(\psi_2\) contributes one level-II and one level-III root,
and \(\phi_2\) is the two-level-II contact state carrying one level-III root.
\end{proof}

The dynamic spin-chain basis contains explicit \(Z^\pm\) markers.  In the
string basis one absorbs half of this marker into each boson:
\[
  \ket{\phi_A(p)}_{\rm str}=\ket{\phi_A(p)(Z^+)^{1/2}},
  \qquad
  \ket{\psi_\alpha(p)}_{\rm str}=\ket{\psi_\alpha(p)} .
\]
The scattering factors which enter
\eqref{eq:planar-n4-single-copy-nested-bethe-yang} are then replaced by
\begin{equation}
\begin{aligned}
  S^{{\rm I},{\rm I}}_{\rm str}(p_1,p_2)
  &=
  S^{{\rm I},{\rm I}}(p_1,p_2)
  \sqrt{\frac{x_1^+x_2^-}{x_1^-x_2^+}},\\
  S^{{\rm II},{\rm I}}_{\rm str}(y,p)
  &=
  \frac{y-x^-(p)}{y-x^+(p)}
  \sqrt{\frac{x^+(p)}{x^-(p)}},\\
  S^{{\rm III},{\rm II}}_{\rm str}(w,y)
  &=
  S^{{\rm III},{\rm II}}(w,y),
  \qquad
  S^{{\rm III},{\rm III}}_{\rm str}(w_1,w_2)
  =
  S^{{\rm III},{\rm III}}(w_1,w_2).
\end{aligned}
  \label{eq:planar-n4-string-basis-nesting-factors}
\end{equation}
For the full \(\mathfrak{su}(2|2)_{\rm c}\otimes
\mathfrak{su}(2|2)_{\rm c}\) magnon, the left and right copies have
independent roots \(y_a^{(\varsigma)},w_b^{(\varsigma)}\),
\(\varsigma=L,R\).  The physical level-I Bethe equation is obtained by
multiplying the left and right auxiliary factors and by using the chosen
physical vacuum scalar factor.  In the \(SU(2)\)-vacuum nesting, the
auxiliary insertion on the \(k\)-th physical magnon is
\[
  \prod_{\varsigma=L,R}\prod_{a=1}^{K^{\rm II}_{(\varsigma)}}
  \frac{y_a^{(\varsigma)}-x_k^-}{y_a^{(\varsigma)}-x_k^+}
  \sqrt{\frac{x_k^+}{x_k^-}}.
\]
In the alternative \(SL(2)\)-vacuum nesting this factor is inverted:
\[
  \prod_{\varsigma=L,R}\prod_{a=1}^{K^{\rm II}_{(\varsigma)}}
  \frac{y_a^{(\varsigma)}-x_k^+}{y_a^{(\varsigma)}-x_k^-}
  \sqrt{\frac{x_k^-}{x_k^+}}.
\]
This reciprocal relation is a nesting-frame statement, not a change of the
BES dressing phase.  Thus the nested Bethe equations are the
diagonalization of a matrix-valued factorized scattering problem, not
additional dynamical assumptions.

\begin{controlledapproximation}[Asymptotic long-chain regime]
\label{ca:planar-n4-asymptotic-regime}
Equation~\eqref{eq:planar-n4-rank-one-stringbook-aba} is an instance of
Assumption~\ref{ass:planar-n4-aba-regime}.  At weak coupling, wrapping
effects begin at the loop order at which planar interactions can close around
the trace.  At finite coupling, the distinction is between the asymptotic
Bethe--Yang problem and the finite-size problem solved by mirror TBA or the
quantum spectral curve.
\end{controlledapproximation}

\section{Weak-Coupling Reduction}

The weak-coupling limit is a useful place to separate two conventions that
are often conflated.  The stringbook rank-one equation
\eqref{eq:planar-n4-rank-one-stringbook-aba} uses the \(SL(2)\)-vacuum
orientation.  The compact \(SU(2)\) coordinate Bethe equation of
Chapter~\ref{ch:planar-n4-spectral-problem-spin-chains} uses the reciprocal
Bethe--Yang phase, because there the scattering phase is the inverse of the
ordered-chamber exchange coefficient.  The two statements are compatible,
but only after the orientation is named.

\paragraph{Weak limit of the rank-one ABA orientation.}
\label{par:planar-n4-rank-one-aba-weak-limit}
Let \(u_j,u_k\) stay away from the shifted branch points and from each other
as \(g\to0\), and take the physical outside branch
\[
  x_j^\pm=x(u_j\pm\ii/2),
  \qquad
  x(U)+x(U)^{-1}=\frac{U}{g}.
\]
Then
\begin{align}
  x_j^\pm
  &=
  \frac{u_j\pm\ii/2}{g}
  -\frac{g}{u_j\pm\ii/2}
  +O(g^3),
  \label{eq:planar-n4-xpm-weak-branch-expansion}\\
  \frac{x_j^+}{x_j^-}
  &=
  \frac{u_j+\ii/2}{u_j-\ii/2}+O(g^2),
  \label{eq:planar-n4-weak-momentum-ratio}\\
  \mathcal R_{jk}
  &:=
  \frac{x_j^- -x_k^+}{x_j^+ -x_k^-}
  \frac{1-1/(x_j^+x_k^-)}{1-1/(x_j^-x_k^+)}
  =
  \frac{u_j-u_k-\ii}{u_j-u_k+\ii}+O(g^2),
  \label{eq:planar-n4-rank-one-rational-weak-limit}\\
  H(u_j)-1
  &=
  2\ii g\left(\frac1{x_j^+}-\frac1{x_j^-}\right)
  =
  \frac{2g^2}{u_j^2+1/4}+O(g^4).
  \label{eq:planar-n4-rank-one-energy-weak-limit}
\end{align}
The dressing phase satisfies \(\sigma^2=1+O(g^6)\) for fixed weak Bethe
roots in the charge-expansion regime.  Hence
\eqref{eq:planar-n4-rank-one-stringbook-aba} reduces to
\begin{equation}
  \left(\frac{u_j+\ii/2}{u_j-\ii/2}\right)^L
  =
  \prod_{\substack{k=1\\k\neq j}}^M
  \frac{u_j-u_k-\ii}{u_j-u_k+\ii},
  \label{eq:planar-n4-rank-one-sl2-weak-bethe}
\end{equation}
the \(SL(2)\)-orientation also used in
\eqref{eq:planar-n4-sl2-one-loop-bethe} and in the weak-QSC Baxter
degeneration of Chapter~\ref{ch:planar-n4-quantum-spectral-curve-hexagon}.
The reciprocal factor \(\mathcal R_{jk}^{-1}\) gives the compact \(SU(2)\)
one-loop phase of \eqref{eq:planar-n4-one-loop-su2-bethe}.

The convention check is an ordinary weak-branch expansion.  The outside branch
of \(x+x^{-1}=U/g\) has the expansion
\[
  x(U)=\frac{U}{g}-\frac{g}{U}+O(g^3),
\]
which gives \eqref{eq:planar-n4-xpm-weak-branch-expansion}.  Dividing the
two shifted expansions proves \eqref{eq:planar-n4-weak-momentum-ratio}.  In
\(\mathcal R_{jk}\), the first factor has numerator
\[
  \frac{u_j-\ii/2}{g}-\frac{u_k+\ii/2}{g}+O(g)
  =
  \frac{u_j-u_k-\ii}{g}+O(g)
\]
and denominator \((u_j-u_k+\ii)/g+O(g)\).  The second factor equals
\(1+O(g^2)\), because each \(x^\pm\) is \(O(g^{-1})\).  This proves
\eqref{eq:planar-n4-rank-one-rational-weak-limit}.  The energy expansion is
\[
  2\ii g^2\left(
    \frac{1}{u_j+\ii/2}-\frac{1}{u_j-\ii/2}
  \right)+O(g^4)
  =
  \frac{2g^2}{u_j^2+1/4}+O(g^4).
\]
Finally, the first weak BES coefficient is \(c_{2,3}(g)=O(g^3)\), while the
weak charges give \(q_2q_3-q_3q_2=O(g^3)\); the dressing phase therefore
cannot affect the one-loop equation.  Substituting the limits into
\eqref{eq:planar-n4-rank-one-stringbook-aba} gives
\eqref{eq:planar-n4-rank-one-sl2-weak-bethe}.

Using \(g^2=\lambda/(16\pi^2)\), the leading energy term in
\eqref{eq:planar-n4-rank-one-energy-weak-limit} agrees with the one-loop
normalization of Chapters~\ref{ch:planar-n4-spectral-problem-spin-chains}
and~\ref{ch:planar-n4-quantum-spectral-curve-hexagon}.

\section{Bound States and the Large-Spin Cusp Limit}

A pole of the \(SL(2)\) scalar factor at \(x_1^-=x_2^+\) gives a Bethe
string.  More precisely, the displayed \(SL(2)\)-vacuum rational factor has
a simple kinematic pole when the outgoing endpoint of the first constituent
equals the incoming endpoint of the second constituent.  A \(Q\)-magnon
string with real center \(u\) is described by
\[
  u_j=u+\frac{\ii}{2}(Q-2j+1),
  \qquad j=1,\ldots,Q,
\]
with constituent variables \(x_j^\pm=x(u_j\pm\ii/2)\).  Put
\begin{equation}
  x_Q^+(u)=x\!\left(u+\frac{\ii Q}{2}\right)=x_1^+,
  \qquad
  x_Q^-(u)=x\!\left(u-\frac{\ii Q}{2}\right)=x_Q^- .
  \label{eq:planar-n4-bound-state-endpoints}
\end{equation}
These are the only Zhukovsky endpoints that survive after fusion.

\begin{proposition}[Fusion of a Bethe string]
\label{prop:planar-n4-bound-state-fusion}
Assume a \(Q\)-string is formed by the adjacent-pole conditions
\[
  x_j^-=x_{j+1}^+,\qquad j=1,\ldots,Q-1,
\]
and that the path to this pole does not cross a dressing-phase branch
singularity.  Then the total momentum and physical energy of the string are
\begin{equation}
  \ee^{\ii p}=\frac{x_1^+}{x_Q^-},
  \qquad
  H_Q(u)=Q+2\ii g\left(\frac1{x_Q^+(u)}-\frac1{x_Q^-(u)}\right),
  \label{eq:planar-n4-bound-state-momentum-energy}
\end{equation}
and the shortening identity gives
\begin{equation}
  H_Q(p)=\sqrt{Q^2+16g^2\sin^2\frac p2}.
  \label{eq:planar-n4-bound-state-dispersion}
\end{equation}
If \(y\) is a level-II auxiliary root in the \(SL(2)\)-vacuum string basis,
then its product of constituent scattering factors telescopes:
\begin{equation}
  \prod_{j=1}^{Q}
  \frac{y-x_j^+}{y-x_j^-}
  \sqrt{\frac{x_j^-}{x_j^+}}
  =
  \frac{y-x_Q^+(u)}{y-x_Q^-(u)}
  \sqrt{\frac{x_Q^-(u)}{x_Q^+(u)}} .
  \label{eq:planar-n4-bound-state-auxiliary-telescope}
\end{equation}
For two bound states of charges \(Q_1,Q_2\) and centers \(u,v\), the fused
\(SL(2)\)-vacuum scalar factor is
\begin{equation}
  S^{Q_1,Q_2}_{SL(2)}(u,v)
  =
  \prod_{m=1}^{Q_1}\prod_{n=1}^{Q_2}
  S_{SL(2)}
  \!\left(
    u+\frac{\ii}{2}(Q_1-2m+1),
    v+\frac{\ii}{2}(Q_2-2n+1)
  \right),
  \label{eq:planar-n4-fused-sl2-bound-state-scalar}
\end{equation}
followed by projection to the short \(Q_1\)- and \(Q_2\)-particle
representations.  The product formula is the scalar part of fusion; the
matrix projection and the analytic continuation of the dressing factor are
separate representation-theoretic and scalar-branch inputs.
\end{proposition}

\begin{proof}
The condition \(u_{j+1}=u_j-\ii\) gives
\[
  x_j^-=x(u_j-\ii/2)=x(u_{j+1}+\ii/2)=x_{j+1}^+,
\]
so all internal endpoints cancel in products.  Hence
\[
  \prod_{j=1}^{Q}\frac{x_j^+}{x_j^-}
  =
  \frac{x_1^+}{x_Q^-}
  =
  \frac{x_Q^+(u)}{x_Q^-(u)}.
\]
The same cancellation in the sum of the physical-sheet one-particle energies
\[
  H_j=1+2\ii g\left(\frac1{x_j^+}-\frac1{x_j^-}\right)
\]
gives the energy formula in
\eqref{eq:planar-n4-bound-state-momentum-energy}.  The bound-state
shortening relation is the centrally extended one with \(H^2-Q^2=4PK\), and
the total central charges telescope to
\[
  P=g(1-\ee^{\ii p}),
  \qquad
  K=g(1-\ee^{-\ii p}).
\]
Therefore
\[
  H_Q^2=Q^2+4g^2(1-\ee^{\ii p})(1-\ee^{-\ii p})
  =
  Q^2+16g^2\sin^2\frac p2 .
\]
Taking the positive physical branch gives
\eqref{eq:planar-n4-bound-state-dispersion}.

The auxiliary factor
\eqref{eq:planar-n4-bound-state-auxiliary-telescope} is the same endpoint
cancellation:
\[
  \prod_{j=1}^{Q}\frac{y-x_j^+}{y-x_j^-}
  =
  \frac{y-x_1^+}{y-x_Q^-},
  \qquad
  \prod_{j=1}^{Q}\frac{x_j^-}{x_j^+}
  =
  \frac{x_Q^-}{x_1^+}.
\]
Finally, scalar fusion means that a distant probe scatters consecutively
through every constituent of each string before the internal tensor product is
projected to the short representation.  This gives
\eqref{eq:planar-n4-fused-sl2-bound-state-scalar}.  Internal pole residues
and the matrix projection determine the normalization of the bound-state
intertwiner; they are not additional endpoint identities and are kept as
fusion inputs here.
\end{proof}

The resulting asymptotic Bethe--Yang equations for \(K^{\rm I}\) bound-state
particles of charges \(Q_k\) are obtained from
\eqref{eq:planar-n4-single-copy-nested-bethe-yang} by replacing the
fundamental scalar factor with
\eqref{eq:planar-n4-fused-sl2-bound-state-scalar} and the fundamental
auxiliary factors with
\eqref{eq:planar-n4-bound-state-auxiliary-telescope}.  Thus the first
equation in the \(SL(2)\)-vacuum frame reads
\begin{equation}
  \ee^{\ii p_kL}
  =
  \prod_{\substack{j=1\\j\neq k}}^{K^{\rm I}}
  S^{Q_j,Q_k}_{SL(2)}(u_j,u_k)
  \prod_{\varsigma=L,R}
  \prod_{\ell=1}^{K^{\rm II}_{(\varsigma)}}
  \frac{y_\ell^{(\varsigma)}-x_{Q_k}^+(u_k)}
       {y_\ell^{(\varsigma)}-x_{Q_k}^-(u_k)}
  \sqrt{\frac{x_{Q_k}^-(u_k)}{x_{Q_k}^+(u_k)}} .
  \label{eq:planar-n4-bound-state-bethe-yang}
\end{equation}
The level-II and level-III equations keep the same rational form as before,
with each fundamental \(x^\pm\) replaced by the appropriate bound-state
endpoints \(x_Q^\pm(u)\).  This is the bound-state input used by the mirror
TBA node \(Q\mid\bullet\) in
Chapter~\ref{ch:planar-n4-mirror-tba-y-system}.

In the \(SL(2)\) sector, twist-\(L\) spin-\(S\) operators are
\[
  \operatorname{tr}(D_+^{s_1}Z\cdots D_+^{s_L}Z),
  \qquad \sum_{j=1}^Ls_j=S .
\]
The one-loop Bethe equations in this noncompact sector are
\begin{equation}
  \left(\frac{u_k+\ii/2}{u_k-\ii/2}\right)^L
  =
  \prod_{\substack{j=1\\j\neq k}}^S
  \frac{u_k-u_j-\ii}{u_k-u_j+\ii},
  \qquad k=1,\ldots,S,
  \label{eq:planar-n4-sl2-one-loop-bethe}
\end{equation}
with cyclicity selecting zero total spin-chain momentum.  For the minimal
large-spin trajectory the roots are real, symmetric under \(u\mapsto -u\),
and have mode numbers \(\pm1\) in the logarithmic equation.  Write
\[
  u_k=S\bar u_k,
  \qquad
  \int_{-a}^a\bar\rho_0(\bar u)\,\dd\bar u=1,
\]
where \(\bar\rho_0\) is the continuum density of the scaled roots.  In the
limit \(S\gg L\), the derivative of the continuum logarithmic equation is
encoded by the resolvent
\[
  G_0(z)=\int_{-a}^a\frac{\bar\rho_0(v)}{z-v}\,\dd v,
  \qquad
  G_0(z)=\frac1z+O(z^{-2}).
\]
The discontinuity convention is
\[
  G_0(y+\ii0)-G_0(y-\ii0)=-2\pi\ii\,\bar\rho_0(y).
\]
After differentiating the continuum Bethe equation, the boundary values obey
\[
  G_0'(y+\ii0)+G_0'(y-\ii0)=4\pi\delta(y)
  \qquad (-a<y<a).
\]
The unique solution with one symmetric cut and the displayed normalization is
\begin{equation}
  G_0(z)
  =
  -\ii
  \log\frac{\sqrt{4z^2-1}+\ii}{\sqrt{4z^2-1}-\ii},
  \qquad a=\frac12,
  \label{eq:planar-n4-sl2-cusp-resolvent}
\end{equation}
where the square root is chosen so that \(\sqrt{4z^2-1}\sim2z\) at infinity.
Its density is
\begin{equation}
  \bar\rho_0(y)
  =
  \frac1\pi
  \log
  \frac{1+\sqrt{1-4y^2}}{1-\sqrt{1-4y^2}},
  \qquad |y|<\frac12 .
  \label{eq:planar-n4-sl2-cusp-density}
\end{equation}
The one-loop anomalous dimension is
\[
  \gamma^{(1)}
  =
  2g^2\sum_{k=1}^S\frac{1}{u_k^2+1/4}.
\]
The singular region near \(u=0\) is resolved by evaluating the resolvent at
\(z=\ii/(2S)\):
\[
  \gamma^{(1)}
  =
  -4g^2\,\operatorname{Im}G_0\!\left(\frac{\ii}{2S}\right)+O(S^0g^2).
\]
Using the continuation
\[
  G_0\!\left(\frac{\ii}{2S}\right)
  =
  -\ii
  \log
  \frac{\sqrt{1+S^{-2}}+1}{\sqrt{1+S^{-2}}-1},
\]
one obtains
\begin{equation}
  \Delta-S-L
  =
  \gamma^{(1)}
  =
  8g^2\log S+O(S^0g^2,g^4).
  \label{eq:planar-n4-sl2-one-loop-cusp-log}
\end{equation}

\section{Finite-Density Bethe Roots and Spectral Curves}
\label{sec:planar-n4-finite-density-spectral-curves}

The phrase ``finite gap'' will be used here in a precise Bethe-ansatz sense.
It is the statement that, in a large-charge limit, Bethe roots condense on
finitely many arcs and that the Cauchy transform of the limiting measure
extends to a meromorphic object on a finite-sheeted Riemann surface.  This
surface is the spectral curve of the state.

\begin{definition}[Finite-density Bethe-root curve]
\label{def:finite-density-bethe-root-curve}
Let \(N\) be a scaling parameter and let
\(\{u_{j,N}\}_{j=1}^{M_N}\subset\mathbb C\) be Bethe roots with
\(M_N\to\infty\).  Suppose that, after a scaling
\(u_{j,N}=N\bar u_{j,N}\), the empirical measures
\[
  \mu_N=\frac1{M_N}\sum_{j=1}^{M_N}\delta_{\bar u_{j,N}}
\]
converge weakly to an absolutely continuous measure
\[
  \dd\mu(u)=\rho(u)\,\dd s
\]
supported on a finite union of oriented analytic arcs
\(\mathcal C=\bigcup_{a=1}^r\mathcal C_a\).  The Cauchy transform is
\[
  G(z)=\int_{\mathcal C}\frac{\rho(u)\,\dd s(u)}{z-u},
  \qquad z\in\mathbb C\setminus\mathcal C .
\]
A finite-density Bethe-root curve is the data
\[
  \left(\Sigma,\pi:\Sigma\to\mathbb P^1_z,\lambda,\{A_a,B_a\}\right)
\]
where \(\Sigma\) is a compact Riemann surface obtained by gluing sheets across
the arcs \(\mathcal C_a\), \(\lambda\) is a meromorphic differential whose
physical-sheet primitive contains \(G(z)\dd z\), and the cycles \(A_a\)
encircle the cuts.  The filling fractions are the periods
\[
  \mathcal S_a=\frac{1}{2\pi\ii}\oint_{A_a}\lambda .
\]
\end{definition}

The definition separates three pieces of information that are often blurred.
The measure \(\rho\) is the continuum limit of roots.  The Riemann surface
\(\Sigma\) records analytic continuation across root cuts.  The meromorphic
differential \(\lambda\) records the logarithmic Bethe equation: mode numbers
are period data, not decorations on a drawing of cuts.

\paragraph{Jump data from the Cauchy transform.}
Let \(G\) be the Cauchy transform in
Definition~\ref{def:finite-density-bethe-root-curve}, and assume that
\(\rho\) is Hölder continuous on the interior of a cut \(\mathcal C_a\).
For \(u\in\mathcal C_a^\circ\), denote by \(G_+(u)\) and \(G_-(u)\) the
boundary values from the two sides of the oriented arc.  Then
\begin{equation}
  G_+(u)-G_-(u)=-2\pi\ii\,\rho(u),
  \qquad
  \frac{G_+(u)+G_-(u)}2
  =
  \operatorname{p.v.}\int_{\mathcal C}
  \frac{\rho(v)\,\dd s(v)}{u-v}.
  \label{eq:finite-density-cauchy-jump}
\end{equation}
Consequently, whenever the continuum logarithmic Bethe equation has the form
\begin{equation}
  2\,\operatorname{p.v.}\int_{\mathcal C}
  \frac{\rho(v)\,\dd s(v)}{u-v}
  =
  V'(u)-2\pi n_a,\qquad u\in\mathcal C_a^\circ,
  \label{eq:finite-density-continuum-bethe}
\end{equation}
with integer mode number \(n_a\) on \(\mathcal C_a\), the quasimomentum
\begin{equation}
  p(z)=\frac12V'(z)-G(z)
  \label{eq:finite-density-quasimomentum}
\end{equation}
satisfies the gluing condition
\begin{equation}
  p_+(u)+p_-(u)=2\pi n_a .
  \label{eq:finite-density-quasimomentum-gluing}
\end{equation}

The two boundary-value identities are the Plemelj formulas for the Cauchy
transform on a smooth oriented arc.  Substituting the second identity into
\eqref{eq:finite-density-continuum-bethe} gives
\[
  G_+(u)+G_-(u)=V'(u)-2\pi n_a .
\]
Using \(p_\pm(u)=V'(u)/2-G_\pm(u)\) gives
\[
  p_+(u)+p_-(u)
  =
  V'(u)-\bigl(G_+(u)+G_-(u)\bigr)
  =
  2\pi n_a .
\]

For the large-spin one-loop cusp density above, the spectral curve is already
visible.  Put
\[
  y^2=4z^2-1,
  \qquad
  G_0(z)
  =
  -\ii
  \log\frac{y+\ii}{y-\ii},
  \qquad y\sim2z\quad(z\to\infty).
\]
The two sheets are exchanged by \(y\mapsto -y\).  Under this exchange,
\[
  \exp\!\bigl(\ii G_0\bigr)
  =
  \frac{y+\ii}{y-\ii}
  \quad\longmapsto\quad
  \frac{y-\ii}{y+\ii}
  =
  \exp\!\bigl(-\ii G_0\bigr),
\]
so the cut \([-1/2,1/2]\) is the branch locus of the two-sheeted curve.  The
density \eqref{eq:planar-n4-sl2-cusp-density} is precisely the discontinuity
of this primitive across the cut, and the endpoint square-root behavior is
the local branch structure of \(\Sigma\).

The all-loop ABA refines this large-spin calculation rather than replacing
it.  The macroscopic roots at \(|u|\sim S\) still produce the logarithmic
enhancement, but the all-loop phase induces an \(O(\log S)\) fluctuation of
the density near \(u=O(1)\).  This near-origin fluctuation is the object
described by the BES equation.

\begin{assumption}[Large-spin BES scaling regime]
\label{ass:planar-n4-bes-scaling-regime}
Use the \(SL(2)\) all-loop asymptotic Bethe equation in the regime
\[
  S\to\infty,\qquad L\ \text{fixed},\qquad \log S\gg L,
\]
with symmetric mode numbers \(n_k=\pm1\), zero total momentum, the physical
Zhukovsky branch for the Bethe roots before crossing, and the BES dressing
phase of Definition~\ref{def:planar-n4-dressing-phase-data}.  Assume that the
large-spin root density has a one-cut bulk part at \(u=O(S)\) and a
near-origin correction at \(u=O(1)\) whose coefficient is proportional to
the scaling function.  The all-loop ABA is being used here in precisely the
finite-density sense of Assumption~\ref{ass:planar-n4-aba-regime}; wrapping
effects are outside this asymptotic calculation.
\end{assumption}

The derivative of the logarithmic all-loop \(SL(2)\) Bethe equation has three
types of terms.  The rational XXX kernel gives the one-loop singular source
already solved above.  The rational all-loop Zhukovsky factors and the
dressing phase give additional kernels.  Finally, the \(u=O(S)\) bulk roots
feed a source into the \(u=O(1)\) equation.  After the bulk density has been
used to solve the outer problem, the inner equation is closed for a single
Fourier density.  We use the standard BES normalization
\[
  \widehat\sigma(t)
  =
  \ee^{-t/2}\int_{-\infty}^{\infty}\dd u\,
  \ee^{-\ii t u}\sigma(u),
  \qquad t>0,
\]
where \(\sigma(u)\) denotes the inner density correction in the present
large-spin normalization.

\begin{definition}[BES kernel in the stringbook convention]
\label{def:planar-n4-bes-kernel}
Let \(J_n\) be the Bessel function of the first kind.  Set
\begin{equation}
  K_0(t,t')
  =
  \frac{J_1(t)J_0(t')-J_0(t)J_1(t')}{t-t'},
  \label{eq:planar-n4-bes-rational-kernel}
\end{equation}
with the diagonal value defined by analytic continuation.  The dressing part
is
\begin{equation}
  K_{\rm dr}(t,t';g)
  =
  \frac{4}{tt'}
  \sum_{k,\ell\geq1}
  (-1)^{k+\ell}c_{2k,2\ell+1}(g)
  J_{2\ell}(t)J_{2k-1}(t'),
  \label{eq:planar-n4-bes-dressing-kernel}
\end{equation}
where \(c_{r,s}(g)=-c_{s,r}(g)\) for \(r>s\).  The BES kernel used below is
\[
  K_{\rm BES}(t,t';g)=K_0(t,t')+K_{\rm dr}(t,t';g),
  \qquad
  K_{\rm BES}(t,0;g)
  =
  \frac{J_1(t)}{t}
  +
  \frac{2}{t}\sum_{\ell\geq1}
  (-1)^{\ell-1}c_{2,2\ell+1}(g)J_{2\ell}(t).
\]
\end{definition}

The Fourier transform which turns powers of \(x^\pm(u)^{-1}\) into Bessel
functions is the convention-sensitive step.  We spell it out because replacing
the signed ratio \(J_m(2gt)/t\) by \(J_m(2g|t|)/|t|\) loses a parity sign in
the lower half-plane transform.

\begin{proposition}[Zhukovsky Fourier transform]
\label{prop:planar-n4-bes-fourier-xpm}
Let \(g>0\), let \(m\in\mathbb Z_{>0}\), and let \(x(z)\) be the physical
outside branch of
\[
  x(z)+x(z)^{-1}=\frac{z}{g},
  \qquad x(z)\sim \frac{z}{g}\quad (z\to\infty),
\]
with short cut \([-2g,2g]\).  Put
\[
  x^\pm(u)=x(u\pm \ii/2),
  \qquad u\in\mathbb R .
\]
For \(t\neq0\), with \(\Theta(t)=1\) for \(t>0\) and \(0\) for \(t<0\),
the oscillatory Fourier transform is
\begin{equation}
  \int_{-\infty}^{\infty}\frac{\dd u}{2\pi}\,
  \ee^{-\ii t u}\frac1{x^\pm(u)^m}
  =
  (-\ii)^m m\,\ee^{-|t|/2}\,
  \frac{\pm\Theta(\pm t)J_m(2gt)}{t}.
  \label{eq:planar-n4-bes-fourier-xpm}
\end{equation}
Here the upper sign is used for \(x^+\), the lower sign for \(x^-\), and
\(J_m\) is the Bessel function of the first kind.
\end{proposition}

\begin{proof}
For the \(x^+\) transform set \(z=u+\ii/2\).  Then
\[
  \ee^{-\ii t u}=\ee^{-t/2}\ee^{-\ii t z},
\]
and the integration line is \(\mathbb R+\ii/2\).  If \(t<0\), this line is
closed in the upper half-plane and no cut is crossed, so the integral
vanishes.  If \(t>0\), close in the lower half-plane.  The deformation crosses
the Zhukovsky cut and gives the negatively oriented loop around it.  Using
\(z=g(x+x^{-1})\) on this loop, with \(x\) running around \(|x|=1\), gives
\[
  I_m^+(t)
  =
  -\ee^{-t/2}\oint_{|x|=1}\frac{\dd x}{2\pi}\,
  g(1-x^{-2})x^{-m}\ee^{-\ii gt(x+x^{-1})}.
\]
The generating function
\[
  \ee^{-\ii gt(x+x^{-1})}
  =
  \sum_{n\in\mathbb Z}(-\ii)^nJ_n(2gt)x^n
\]
therefore yields
\[
  I_m^+(t)
  =
  -\ii g\ee^{-t/2}
  \left[(-\ii)^{m-1}J_{m-1}(2gt)
        -(-\ii)^{m+1}J_{m+1}(2gt)\right].
\]
Since \((-\ii)^{m+1}=-(-\ii)^{m-1}\), and
\[
  J_{m-1}(z)+J_{m+1}(z)=\frac{2m}{z}J_m(z),
\]
this becomes
\[
  I_m^+(t)=(-\ii)^m m\,\ee^{-t/2}\frac{J_m(2gt)}{t},
  \qquad t>0.
\]
Together with the vanishing for \(t<0\), this proves the upper-sign formula.

For real \(u\), the lower shifted branch satisfies
\[
  x^-(u)=\overline{x^+(u)}.
\]
Hence
\[
  I_m^-(t)=\overline{I_m^+(-t)}.
\]
Applying the upper-sign result to \(-t>0\) and using
\(J_m(-z)=(-1)^mJ_m(z)\) gives the lower-sign formula in
\eqref{eq:planar-n4-bes-fourier-xpm}.  The factor \(J_m(2gt)/t\), with
signed \(t\), is therefore part of the convention rather than a cosmetic
choice.
\end{proof}

The signs in Proposition~\ref{prop:planar-n4-bes-fourier-xpm} encode the same
\(x^+\)-above, \(x^-\)-below sheet convention as the mirror continuation in
Chapter~\ref{ch:planar-n4-mirror-tba-y-system}.  Reversing them would invert
the dressing-source contribution.

\paragraph{BES equation from the large-spin ABA.}
\label{par:planar-n4-bes-equation}
Under Assumption~\ref{ass:planar-n4-bes-scaling-regime}, and after applying
the Fourier transform \eqref{eq:planar-n4-bes-fourier-xpm}, the inner
large-spin density obeys
\begin{equation}
  \widehat\sigma(t)
  =
  \frac{t}{\ee^t-1}
  \left[
    K_{\rm BES}(2gt,0;g)
    -
    4g^2\int_0^\infty\dd t'\,
    K_{\rm BES}(2gt,2gt';g)\widehat\sigma(t')
  \right],
  \qquad t>0.
  \label{eq:planar-n4-bes-equation}
\end{equation}
The corresponding large-spin scaling function in the normalization of
\eqref{eq:planar-n4-sl2-one-loop-cusp-log} is
\begin{equation}
  \Delta-S-L=f(g)\log S+O(S^0),
  \qquad
  f(g)
  =
  8g^2
  -
  64g^4
  \int_0^\infty\dd t\,
  \widehat\sigma(t)\frac{J_1(2gt)}{2gt}.
  \label{eq:planar-n4-bes-scaling-function}
\end{equation}

We spell out the logic because it is easy to hide a convention change in this
calculation.  Taking the logarithm of the \(SL(2)\) all-loop Bethe equation
and differentiating with respect to the target rapidity \(u\) gives the
rational kernel
\[
  \frac{2}{(u-v)^2+1}-2\pi\delta(u-v),
\]
where the delta term records the branch of the arccotangent scattering phase.
The all-loop rational Zhukovsky factor contributes
\[
  -2\ii\,\partial_u\log
  \frac{1-1/(x^+(u)x^-(v))}
       {1-1/(x^-(u)x^+(v))},
\]
and the scalar factor contributes
\[
  2\partial_u\theta(x^\pm(v),x^\pm(u)).
\]
The \(O(S)\) bulk roots are described by the one-cut density
\eqref{eq:planar-n4-sl2-cusp-density}; expanding their contribution for
fixed \(u\) produces the driving term
\begin{equation}
  \partial_u\left[
  \frac1{2g}\left(\frac1{x^+(u)}+\frac1{x^-(u)}\right)
  +
  \frac{\ii}{2g}
  \sum_{n\geq1}\frac{c_{2,2n+1}(g)}{n}
  \left(\frac1{x^+(u)^{2n}}-\frac1{x^-(u)^{2n}}\right)
  \right].
  \label{eq:planar-n4-bes-inner-driving}
\end{equation}
The explicit \(L\)-dependent terms in the logarithmic Bethe equation are
smaller than this driving term in the scaling regime \(\log S\gg L\), so they
affect the \(O(S^0)\) part rather than \(f(g)\).

Applying \((2\pi)^{-1}\ee^{-t/2}\int\dd u\,\ee^{-\ii t u}\) to the inner
equation gives \eqref{eq:planar-n4-bes-equation}.  The Bessel transform
\eqref{eq:planar-n4-bes-fourier-xpm} turns the all-loop rational factors into
\(K_0\), and turns the charge expansion of the dressing phase into
\(K_{\rm dr}\).  The factor \(t/(\ee^t-1)\) is the inverse of the rational
large-spin kernel in this Fourier normalization.  The energy formula follows
by expanding
\[
  \sum_k\left(
  \frac1{x_k^+}-\frac1{x_k^-}
  \right)
\]
against the same bulk-plus-inner density and applying
\eqref{eq:planar-n4-bes-fourier-xpm} with \(m=1\).  The outer density gives
the one-loop \(8g^2\log S\) term; the inner density gives the integral
correction in \eqref{eq:planar-n4-bes-scaling-function}.

\begin{proposition}[Weak expansion of the BES scaling function]
\label{prop:planar-n4-bes-weak-scaling-function}
The BES equation \eqref{eq:planar-n4-bes-equation} implies
\begin{equation}
  f(g)
  =
  8g^2-\frac{8\pi^2}{3}g^4+\frac{88\pi^4}{45}g^6
  -
  \left(\frac{584\pi^6}{315}+64\zeta(3)^2\right)g^8
  +O(g^{10}).
  \label{eq:planar-n4-bes-weak-scaling-function}
\end{equation}
The first dressing-phase contribution enters at order \(g^8\) and supplies
the \(\zeta(3)^2\) term in this normalization.
\end{proposition}

\begin{proof}
Write
\[
  \widehat\sigma(t)=\sigma_0(t)+g^2\sigma_1(t)+g^4\sigma_2(t)+O(g^6).
\]
Since
\[
  \frac{J_1(2gt)}{2gt}
  =
  \frac12-\frac{g^2t^2}{4}+\frac{g^4t^4}{24}+O(g^6),
  \qquad
  K_0(2gt,0)=\frac12-\frac{g^2t^2}{4}+\frac{g^4t^4}{24}+O(g^6),
\]
and
\[
  K_0(2gt,2gt')
  =
  \frac12
  -
  \frac{g^2}{4}(t^2-tt'+t'^2)
  +O(g^4),
\]
the first two density coefficients are
\[
  \sigma_0(t)=\frac{t}{2(\ee^t-1)}
\]
and
\[
  \sigma_1(t)
  =
  \frac{t}{\ee^t-1}
  \left[
    -\frac{t^2}{4}
    -4\int_0^\infty\dd t'\,\frac12\,\sigma_0(t')
  \right]
  =
  -\frac{t}{\ee^t-1}
  \left(\frac{t^2}{4}+\frac{\pi^2}{6}\right),
\]
because
\[
  \int_0^\infty\frac{t\,\dd t}{\ee^t-1}=\frac{\pi^2}{6}.
\]
For the \(g^4\) density coefficient the dressing source is now required.
Using
\[
  c_{2,3}(g)=4\zeta(3)g^3+O(g^5),
  \qquad
  J_2(2gt)=\frac{g^2t^2}{2}+O(g^4),
\]
and the prefactor \(2/(2gt)\) in \(K_{\rm BES}(2gt,0;g)\), one obtains
\[
  K_{\rm dr}(2gt,0;g)
  =
  2\zeta(3)g^4t+O(g^6).
\]
The \(g^4\) part of the BES equation is therefore
\[
\begin{aligned}
  \sigma_2(t)
  &=
  \frac{t}{\ee^t-1}
  \left[
    \frac{t^4}{24}+2\zeta(3)t
    -4\int_0^\infty\dd t'\,
    \left(
      \frac12\sigma_1(t')
      -\frac{t^2-tt'+t'^2}{4}\sigma_0(t')
    \right)
  \right].
\end{aligned}
\]
The required moments are
\[
\begin{gathered}
  \int_0^\infty\sigma_0(t')\,\dd t'=\frac{\pi^2}{12},
  \qquad
  \int_0^\infty t'\sigma_0(t')\,\dd t'=\zeta(3),
  \qquad
  \int_0^\infty t'^2\sigma_0(t')\,\dd t'=\frac{\pi^4}{30},\\
  \int_0^\infty\sigma_1(t')\,\dd t'=-\frac{2\pi^4}{45}.
\end{gathered}
\]
Substitution gives the explicit coefficient
\[
  \sigma_2(t)
  =
  \frac{t}{\ee^t-1}
  \left(
    \frac{t^4}{24}
    +\frac{\pi^2t^2}{12}
    +\zeta(3)t
    +\frac{11\pi^4}{90}
  \right).
\]
Put
\[
  A(g)=
  \int_0^\infty\dd t\,
  \widehat\sigma(t)\frac{J_1(2gt)}{2gt}
  =
  A_0+g^2A_1+g^4A_2+O(g^6).
\]
The two coefficients are
\[
  A_0
  =
  \int_0^\infty\frac{t\,\dd t}{4(\ee^t-1)}
  =
  \frac{\pi^2}{24},
\]
and
\[
  A_1
  =
  \int_0^\infty\frac{t\,\dd t}{\ee^t-1}
  \left(-\frac{t^2}{4}-\frac{\pi^2}{12}\right)
  =
  -\frac{\pi^4}{60}-\frac{\pi^4}{72}
  =
  -\frac{11\pi^4}{360},
\]
using
\[
  \int_0^\infty\frac{t^3\,\dd t}{\ee^t-1}=\frac{\pi^4}{15}.
\]
The next coefficient is
\[
  A_2
  =
  \int_0^\infty\dd t\,
  \left(
    \frac12\sigma_2(t)
    -\frac{t^2}{4}\sigma_1(t)
    +\frac{t^4}{24}\sigma_0(t)
  \right).
\]
Using
\[
  \int_0^\infty\frac{t^2\,\dd t}{\ee^t-1}=2\zeta(3),
  \qquad
  \int_0^\infty\frac{t^5\,\dd t}{\ee^t-1}
  =
  \frac{8\pi^6}{63},
\]
one finds
\[
  A_2=\frac{73\pi^6}{2520}+\zeta(3)^2.
\]
Substitution into
\eqref{eq:planar-n4-bes-scaling-function} gives
\[
  f(g)=8g^2-64g^4A_0-64g^6A_1-64g^8A_2+O(g^{10}),
\]
which is \eqref{eq:planar-n4-bes-weak-scaling-function}.
\end{proof}

\begin{quotedtheorem}[Strong solution of the BES equation]
\label{qt:planar-n4-bes-strong-solution}
Let \(\Gamma_{\rm BES}(g)\) denote the conventional BES cusp function whose
weak expansion begins \(\Gamma_{\rm BES}(g)=4g^2+O(g^4)\).  Define the
Dirichlet beta values
\[
  \beta_{\rm D}(s)=\sum_{n=0}^{\infty}\frac{(-1)^n}{(2n+1)^s},
  \qquad
  \mathrm K=\beta_{\rm D}(2),
\]
where \(\mathrm K\) is Catalan's constant, and set
\[
  c_1=\frac{3\log2}{4\pi},\qquad
  c_2=\frac{\mathrm K}{16\pi^2},\qquad
  c_3=\frac{27\zeta(3)}{2^{11}\pi^3},\qquad
  c_4=\frac{21\,\beta_{\rm D}(4)}{2^{10}\pi^4}.
\]
The strong-coupling solution of the BES Riemann--Hilbert problem gives, as a
Poincare asymptotic expansion on the positive real \(g\)-axis,
\begin{equation}
  \Gamma_{\rm BES}(g)
  =
  2g_{\rm s}
  \left[
    1-\frac{c_2}{g_{\rm s}^2}
    -\frac{c_3}{g_{\rm s}^3}
    -\frac{c_4+2c_2^2}{g_{\rm s}^4}
    +O(g_{\rm s}^{-5})
  \right],
  \qquad
  g_{\rm s}=g-c_1 .
  \label{eq:planar-n4-bes-quoted-strong-gamma}
\end{equation}
The quoted theorem is a theorem about the analytic solution of the BES
integral equation.  Its use in the spectral problem is conditional on the
large-spin BES bridge of
Assumption~\ref{ass:planar-n4-bes-scaling-regime}.
\end{quotedtheorem}

\begin{remark}[Strong BES expansion in the monograph normalization]
\label{prop:planar-n4-bes-strong-monograph-normalization}
With \(f(g)\) defined by
\[
  \Delta-S-L=f(g)\log S+O(S^0)
\]
in \eqref{eq:planar-n4-bes-scaling-function}, the strong-coupling expansion
corresponding to
Quoted Theorem~\ref{qt:planar-n4-bes-strong-solution} is
\begin{equation}
  f(g)
  =
  4g_{\rm s}
  -\frac{4c_2}{g_{\rm s}}
  -\frac{4c_3}{g_{\rm s}^2}
  -\frac{4(c_4+2c_2^2)}{g_{\rm s}^3}
  +O(g_{\rm s}^{-4}),
  \qquad
  g_{\rm s}=g-\frac{3\log2}{4\pi}.
  \label{eq:planar-n4-bes-strong-shifted-f}
\end{equation}
In particular, expanded in the unshifted coupling,
\begin{equation}
  f(g)
  =
  4g-\frac{3\log2}{\pi}
  -\frac{\mathrm K}{4\pi^2g}
  +O(g^{-2}).
  \label{eq:planar-n4-bes-strong-unshifted-f}
\end{equation}
The weak expansion in
Proposition~\ref{prop:planar-n4-bes-weak-scaling-function} begins
\[
  f(g)=8g^2+O(g^4),
\]
whereas \(\Gamma_{\rm BES}(g)=4g^2+O(g^4)\).  Therefore the present
normalization is
\[
  f(g)=2\Gamma_{\rm BES}(g).
\]
Multiplying \eqref{eq:planar-n4-bes-quoted-strong-gamma} by \(2\) gives
\eqref{eq:planar-n4-bes-strong-shifted-f}.  The first two unshifted terms are
obtained from \(4g_{\rm s}=4g-4c_1\), and
\[
  4c_1=\frac{3\log2}{\pi},\qquad
  4c_2=\frac{\mathrm K}{4\pi^2}.
\]
All further displayed terms are \(O(g^{-2})\) or smaller in the unshifted
coupling, giving \eqref{eq:planar-n4-bes-strong-unshifted-f}.
\end{remark}

This is the convention in which the large-spin anomalous dimension has
coefficient \(f(g)\).  Some Wilson-line literature denotes by
\(\Gamma_{\rm cusp}\) half of this quantity; Chapter~\ref{ch:planar-n4-quantum-spectral-curve-hexagon}
states its Wilson-line convention at the point where cusped Wilson loops are
introduced.  The equality between the large-spin scaling function and the
Wilson-line cusp observable is a structural matching within planar
\(\mathcal N=4\) theory; a nonperturbative construction-level proof remains
outside the present monograph.  The QCD counterpart of the same
large-spin/Wilson-line mechanism was defined from a regulated cusped
Wilson-line operator in
Section~\ref{sec:qcd-cusped-wilson-lines-large-spin}: there the lightlike
coefficient \(\Gamma_{\rm cusp}^q\) is a perturbative anomalous dimension
entering DIS endpoint evolution as
\(P_{{\rm ns},N}=-\Gamma_{\rm cusp}^q\log N+O(N^0)\).  The comparison fixes the
shared operator meaning of the word ``cusp'' while keeping the theories'
logical status distinct: QCD uses factorization and perturbative
renormalization of Wilson-line composites, whereas the present planar
\(\mathcal N=4\) construction uses the integrable all-loop spectral problem.
