\chapter{Four-Dimensional N=2 Gauge Dynamics and Seiberg-Witten Theory}
\label{ch:susy-four-dimensional-n2-seiberg-witten}

\ConventionsLorentzian

This chapter develops the first layer of four-dimensional \(\mathcal N=2\)
gauge dynamics.  The emphasis is the Wilsonian low-energy description on the
Coulomb branch, where the exact two-derivative Abelian action is encoded by a
holomorphic prepotential and by periods of a family of curves.

\begin{hypothesis}[Seiberg--Witten Coulomb-branch structure]
\label{hyp:seiberg-witten-coulomb-branch-datum}
A Seiberg--Witten Coulomb-branch structure of rank \(r\) is the following
collection attached to an assumed four-dimensional \(\mathcal N=2\) local
QFT \(\mathfrak T\).
\begin{enumerate}[label=\textnormal{(\roman*)}]
\item A connected complex analytic vacuum branch
  \(\mathcal B\) and a discriminant locus
  \(\Delta\subset\mathcal B\) such that the low-energy theory over
  \(\mathcal B^\circ=\mathcal B\setminus\Delta\) contains \(r\) Abelian
  vector multiplets and no additional massless charged hypermultiplets.
\item A locally constant electromagnetic charge lattice
  \(\Gamma\to\mathcal B^\circ\) with a nondegenerate integral
  antisymmetric Dirac pairing
  \(\langle\cdot,\cdot\rangle:\Gamma\otimes\Gamma\to\mathbb Z\).  A local electric
  polarization is a symplectic frame
  \((\alpha^I,\beta_I)_{I=1}^r\) of \(\Gamma\).
\item A holomorphic period section
  \[
    \Pi=(a_{D,I},a^I)
  \]
  in the dual local system, with monodromy in
  \(Sp(2r,\mathbb Z)\), such that
  \[
    Z_\gamma=\langle\gamma,\Pi\rangle+\sum_\alpha s_\alpha m_\alpha
  \]
  is the central charge of a state of electromagnetic charge
  \(\gamma\) and flavor charges \(s_\alpha\).
\item On each simply connected electric patch, a holomorphic prepotential
  \(\mathcal F(a)\) satisfying
  \[
    a_{D,I}=\frac{\partial\mathcal F}{\partial a^I},\qquad
    \tau_{IJ}=\frac{\partial^2\mathcal F}
      {\partial a^I\partial a^J},\qquad
    \operatorname{Im}\tau>0 ,
  \]
  and the corresponding two-derivative Abelian Wilsonian action with the
  superspace normalization used in this volume.
\item Local singularity data along \(\Delta\): the primitive charges of
  massless BPS states, the associated Picard--Lefschetz monodromies, and the
  convention for the BPS inequality
  \(M_\gamma\ge\sqrt2\,|Z_\gamma|\).
\item When a curve realization is used, a holomorphic family of compact
  Riemann surfaces \(\Sigma_u\), meromorphic differentials
  \(\lambda_{\mathrm{SW}}\), and symplectic one-cycle frames whose periods
  reproduce the section \(\Pi\).  This curve realization is auxiliary data
  for computing the special geometry; it is not by itself a construction of
  the four-dimensional QFT \(\mathfrak T\).
\end{enumerate}
All dynamical conclusions in this chapter are conditional on the existence of
\(\mathfrak T\) and of such a structure.  The open problem of constructing
\(\mathfrak T\) from a regulator and deriving these structures as theorems of QFT
is recorded in Open Problem~\ref{op:theorem-level-seiberg-witten-construction}.
\end{hypothesis}

\begin{definition}[Pure \(\mathfrak{su}(2)\) global and line datum]
\label{def:pure-su2-sw-global-line-datum}
A pure \(\mathfrak{su}(2)\) Seiberg--Witten microscopic datum is
\[
  \mathfrak S_\ell
  =
  \bigl(
    \mathfrak{su}(2),\ \ell,\ \theta\hbox{-periodicity},\
    \mathcal L_\ell,\ u=\langle\operatorname{tr}\phi^2\rangle
  \bigr),
\]
where the local fields are the \(\mathcal N=2\) vector multiplet in the adjoint
representation, and \(\ell\) is one of the absolute line choices
\[
  \ell\in\{SU(2),SO(3)_+,SO(3)_-\}.
\]
Modulo screening by adjoint local particles, the center-valued Wilson--'t Hooft
line classes form \(\mathbb Z_2^{\rm e}\oplus\mathbb Z_2^{\rm m}\) with the
mod-two Dirac pairing.  The three standard maximal isotropic choices are
\[
  \mathcal L_{SU(2)}=\langle e\rangle,\qquad
  \mathcal L_{SO(3)_+}=\langle m\rangle,\qquad
  \mathcal L_{SO(3)_-}=\langle e+m\rangle .
\]
For the simply connected \(SU(2)\) theory, \(\theta\sim\theta+2\pi\).  For a
fixed \(SO(3)_\pm\) absolute theory, \(\theta\sim\theta+4\pi\), while
\(\theta\mapsto\theta+2\pi\) exchanges \(SO(3)_+\) and \(SO(3)_-\) by the
Witten effect.  The period and BPS-particle formulas below use the
root-normalized local particle lattice
\[
  \Gamma_{\rm part}=\mathbb Z\gamma_m\oplus\mathbb Z\gamma_W,\qquad
  \gamma_m=(1,0),\quad \gamma_W=(0,1),
\]
so the massive \(W\)-boson has charge \(\gamma_W\).  This lattice records
local particle charges and central charges.  The full absolute Wilson--'t Hooft
line lattice is the separate datum \(\mathcal L_\ell\) above.
\end{definition}

\begin{conjecture}[Central conjecture: pure \(\mathfrak{su}(2)\) Seiberg--Witten solution]
\label{conj:pure-su2-sw-low-energy-solution}
Fix a datum \(\mathfrak S_\ell\) from
Definition~\ref{def:pure-su2-sw-global-line-datum}.  There exists a
four-dimensional \(\mathcal N=2\) local QFT with that microscopic gauge algebra,
global line datum, theta periodicity, and local chiral coordinate \(u\), whose
Coulomb branch is
\[
  \mathcal B=\mathbb C_u,\qquad
  \Delta=\{\Lambda^2,-\Lambda^2\},
\]
and whose two-derivative Abelian Wilsonian theory over compact subsets of
\(\mathcal B^\circ=\mathcal B\setminus\Delta\) is encoded by the curve
\[
  y^2=(x^2-\Lambda^4)(x-u)
\]
and a Seiberg--Witten differential with periods
\(\Pi=(a_D,a)^T\).  In the root-normalized local particle lattice
\(\Gamma_{\rm part}\), the central charge is
\[
  Z_{(n_m,n_e)}=n_m a_D+n_e a,
\]
the singularities at \(u=\Lambda^2\) and \(u=-\Lambda^2\) contain light
hypermultiplets of charges \((1,0)\) and \((1,-1)\), respectively, and the
protected observables in the two-derivative sector, including the instanton
prepotential coefficients after scheme matching, agree with this period
datum.
\end{conjecture}

The conjecture is the QFT identification.  The special-geometry calculations
below are exact once the curve, differential, periods, and charge lattice have
been supplied; the finite monodromy products and instanton-coefficient matches
are consistency checks of Conjecture~\ref{conj:pure-su2-sw-low-energy-solution},
not a construction of the four-dimensional continuum QFT.

\section{Coulomb-Branch Coordinates}

Let \(G\) be a compact connected gauge group with Lie algebra \(\mathfrak g\).
An \(\mathcal N=2\) vector multiplet contains a gauge field, two gauginos, and
a complex scalar \(\phi\in\mathfrak g_{\C}\).  A Coulomb-branch point is a
vacuum for which \(\phi\) has expectation value in a Cartan subalgebra and
the low-energy gauge group contains an Abelian factor \(U(1)^r\), where
\(r=\operatorname{rank}G\).  Gauge-invariant holomorphic coordinates are
polynomial functions of \(\phi\), for example
\[
  u_k=\left\langle \operatorname{tr}(\phi^k)\right\rangle .
\]
In the notation of Hypothesis~\ref{hyp:seiberg-witten-coulomb-branch-datum},
these \(u_k\) are local coordinates on the branch \(\mathcal B\) only after
one has chosen the microscopic gauge theory and restricted to the component
where the Abelian rank is \(r\).
The trace convention is the convention fixed in the gauge chapters: the
Yang--Mills kinetic term uses
\[
  \frac{1}{4g^2}\operatorname{tr}(F_{\mu\nu}F^{\mu\nu})
\]
with the chosen invariant bilinear form.

\section{Special Coordinates And Prepotential}

On a smooth patch of \(\mathcal B^\circ\), choose electric special coordinates
\[
  a^I,\qquad I=1,\ldots,r .
\]
The two-derivative Abelian Wilsonian action in
Hypothesis~\ref{hyp:seiberg-witten-coulomb-branch-datum} is determined by a
holomorphic function \(\mathcal F(a)\), the prepotential, through
\[
  a_{D,I}=\frac{\partial\mathcal F}{\partial a^I},
  \qquad
  \tau_{IJ}(a)=\frac{\partial^2\mathcal F}{\partial a^I\partial a^J}.
\]
The imaginary part of \(\tau_{IJ}\) is the positive gauge-coupling matrix in
the region where the Abelian Wilsonian description is valid.  Electric-magnetic
duality acts on the column vector
\[
  \Pi=\begin{pmatrix} a_D \\ a \end{pmatrix}
\]
by integral symplectic matrices.  If \(\gamma=(n_m^I,n_{e,I})\) is an
electromagnetic charge vector, its central charge is
\[
  Z_\gamma=n_{e,I}a^I+n_m^I a_{D,I}+\sum_\alpha s_\alpha m_\alpha ,
\]
where \(m_\alpha\) are background flavor masses and \(s_\alpha\) are flavor
charges.  The BPS inequality has the convention
\[
  M_\gamma\geq \sqrt2\,|Z_\gamma|.
\]
States saturating this bound form shortened \(\mathcal N=2\) multiplets.

In \(\mathcal N=1\) language the same local Abelian \(\mathcal N=2\)
Wilsonian vector multiplets are represented by vector superfields \(V^I\)
and chiral superfields \(\Phi^I\) whose scalar components are \(a^I\).  Once
the two-derivative \(\mathcal N=2\) action is assumed to exist, its scalar
metric and gauge kinetic coefficients are not independent data.  They are
encoded by the same holomorphic prepotential:
\begin{equation}
  K_{\mathcal F}(a,\overline a)
  =
  \frac1{2\pi}\operatorname{Im}
  \left(\overline a^{\,I}\frac{\partial\mathcal F}{\partial a^I}\right),
  \qquad
  \tau_{IJ}(a)=\frac{\partial^2\mathcal F}{\partial a^I\partial a^J}.
  \label{eq:rigid-special-kahler-potential}
\end{equation}
The factor \(1/(2\pi)\) is the superspace normalization used in this volume.
It can be absorbed into the definition of the Abelian fields, but it must not
be silently dropped when comparing instanton, period, and component-action
conventions.

\paragraph{Rigid special-K\"ahler metric from the prepotential.}
Let \(U\subset\mathbb C^r\) be a coordinate patch with holomorphic
coordinates \(a^I\), and let \(\mathcal F\) be holomorphic on \(U\).  Define
\(K_{\mathcal F}\) by \eqref{eq:rigid-special-kahler-potential}.  Then
\begin{equation}
  g_{I\bar J}:=
  \frac{\partial^2K_{\mathcal F}}{\partial a^I\partial\overline a^{\,J}}
  =
  \frac1{2\pi}\operatorname{Im}\tau_{IJ}.
  \label{eq:rigid-special-kahler-metric}
\end{equation}
Consequently the scalar kinetic metric is positive exactly on those
Wilsonian patches where \(\operatorname{Im}\tau_{IJ}\) is positive definite.
Linear terms in \(\mathcal F\) change \(K_{\mathcal F}\) by a K\"ahler
transformation, while real quadratic terms shift only \(\operatorname{Re}\tau\)
and hence only the theta-angle part of the Abelian gauge action.
Write
\[
  K_{\mathcal F}
  =
  \frac1{4\pi\ii}
  \left(
    \overline a^{\,L}\mathcal F_L(a)
    -
    a^L\overline{\mathcal F}_L(\overline a)
  \right),
  \qquad
  \mathcal F_L=\frac{\partial\mathcal F}{\partial a^L}.
\]
Treating \(a^I\) and \(\overline a^{\,J}\) as independent complex
coordinates,
\[
  \frac{\partial K_{\mathcal F}}{\partial a^I}
  =
  \frac1{4\pi\ii}
  \left(
    \overline a^{\,L}\mathcal F_{IL}
    -
    \overline{\mathcal F}_I
  \right).
\]
Differentiating once more gives
\[
  \frac{\partial^2 K_{\mathcal F}}
       {\partial a^I\partial\overline a^{\,J}}
  =
  \frac1{4\pi\ii}
  \left(
    \mathcal F_{IJ}
    -
    \overline{\mathcal F}_{IJ}
  \right)
  =
  \frac1{2\pi}\operatorname{Im}\mathcal F_{IJ}.
\]
Since \(\tau_{IJ}=\mathcal F_{IJ}\), this proves
\eqref{eq:rigid-special-kahler-metric}.  If
\(\mathcal F\mapsto \mathcal F+c_Ia^I+c_0\), then
\(K_{\mathcal F}\) changes by
\((2\pi)^{-1}\operatorname{Im}(\overline a^{\,I}c_I)\), a sum of a
holomorphic and an antiholomorphic function, so \(g_{I\bar J}\) is unchanged.
If \(\mathcal F\mapsto \mathcal F+\frac12 q_{IJ}a^Ia^J\) with \(q_{IJ}\)
real and symmetric, then \(\operatorname{Im}\tau\) is unchanged and
\(\operatorname{Re}\tau\) shifts by \(q_{IJ}\).

\section{Seiberg-Witten Period Geometry}

The curve realization in
Hypothesis~\ref{hyp:seiberg-witten-coulomb-branch-datum} consists of a family
of compact Riemann surfaces \(\Sigma_u\), a meromorphic differential
\(\lambda_{\mathrm{SW}}\), and a symplectic basis of one-cycles
\((A_I,B^I)\) such that
\[
  a^I(u)=\frac{1}{2\pi}\oint_{A_I}\lambda_{\mathrm{SW}},
  \qquad
  a_{D,I}(u)=\frac{1}{2\pi}\oint_{B^I}\lambda_{\mathrm{SW}} .
\]
The differential is fixed up to exact terms by requiring that
\(\partial\lambda_{\mathrm{SW}}/\partial u_k\) span holomorphic differentials
on \(\Sigma_u\) and that the large-field asymptotics reproduce the
one-loop Wilsonian coupling.

The period coupling is integrable on a contractible Coulomb-branch patch
under the following geometric hypotheses.  Let \(U\) be a patch on which the
\(a^I\), \(I=1,\ldots,r\), are holomorphic coordinates.  Suppose
\(\pi:\Sigma\to U\) is a holomorphic family of compact Riemann surfaces with a
flat symplectic homology frame \((A_I,B^I)\), and suppose that the
Seiberg-Witten differential \(\lambda_{\mathrm{SW}}\) is chosen so that
\[
  \omega_J:=\frac{\partial\lambda_{\mathrm{SW}}}{\partial a^J}
\]
is a holomorphic one-form on each fiber, with no residue contribution in the
Riemann bilinear relation, and is normalized by
\[
  \frac1{2\pi}\oint_{A_I}\omega_J=\delta_{IJ}.
\]
The period matrix
\[
  \tau_{IJ}
  :=
  \frac{\partial a_{D,I}}{\partial a^J}
  =
  \frac1{2\pi}\oint_{B^I}\omega_J
\]
is symmetric.  Consequently there is a holomorphic prepotential
\(\mathcal F\) on \(U\), unique up to an additive constant, such that
\[
  a_{D,I}=\frac{\partial\mathcal F}{\partial a^I}.
\]
Because the homology frame is flat, differentiating a period at fixed cycle
gives the period of the differentiated differential:
\[
  \frac{\partial a_{D,I}}{\partial a^J}
  =
  \frac1{2\pi}\oint_{B^I}\omega_J .
\]
Apply the Riemann bilinear relation on a fiber to the two holomorphic
one-forms \(\omega_I\) and \(\omega_J\):
\[
  0
  =
  \sum_{K=1}^r
  \left(
    \oint_{A_K}\omega_I\,\oint_{B^K}\omega_J
    -
    \oint_{B^K}\omega_I\,\oint_{A_K}\omega_J
  \right).
\]
The assumed \(A\)-period normalization gives
\[
  0
  =
  2\pi\oint_{B^I}\omega_J
  -
  2\pi\oint_{B^J}\omega_I,
\]
hence \(\tau_{IJ}=\tau_{JI}\).  The holomorphic one-form
\[
  \alpha=\sum_I a_{D,I}(a)\,\dd a^I
\]
therefore obeys
\[
  \dd\alpha
  =
  \sum_{I<J}
  \left(
    \frac{\partial a_{D,J}}{\partial a^I}
    -
    \frac{\partial a_{D,I}}{\partial a^J}
  \right)
  \dd a^I\wedge\dd a^J
  =
  0 .
\]
Since \(U\) is contractible, the ordinary Poincare lemma gives a holomorphic
function
\[
  \mathcal F(a)=\int_{a_\ast}^{a}\alpha
\]
whose value is independent of the path in \(U\).  Then
\(\dd\mathcal F=\alpha\), which is exactly
\(a_{D,I}=\partial\mathcal F/\partial a^I\).  The additive constant depends
on the chosen base point \(a_\ast\) and has no physical content.

\section{Argument Types And Charge-Monodromy Conventions}
\label{sec:sw-argument-types-charge-monodromy}

The Seiberg-Witten construction mixes several kinds of reasoning.  In this
chapter every load-bearing step is to be read with one of the following status
labels.

\begin{definition}[Argument-status labels for Seiberg-Witten theory]
\label{def:sw-argument-status-labels}
A statement in the rank-one Seiberg-Witten construction has one of these
statuses.
\begin{enumerate}[label=\textnormal{(\roman*)}]
\item \emph{Central conjecture}: the exact identification of a supplied
  period datum with the low-energy Wilsonian physics of a specified
  four-dimensional continuum QFT and absolute line datum.
\item \emph{QFT-hypothesis}: an assumption about the existence of the
  four-dimensional continuum theory, its Coulomb-branch vacua, or its
  Wilsonian effective action.
\item \emph{Symmetry-derived}: a consequence of the microscopic global
  symmetries, anomalies, Weyl group, and dimensional analysis.
\item \emph{Holomorphy-derived}: a consequence of holomorphic dependence on
  Coulomb parameters and couplings once the Wilsonian \(\mathcal N=2\)
  framework has been assumed.
\item \emph{Special-geometry theorem}: a statement about periods, symplectic
  lattices, Riemann bilinear identities, or nodal degenerations of curves.
\item \emph{Constraint-derived}: the minimal object satisfying the preceding
  symmetry, monodromy, and asymptotic constraints, with uniqueness only inside
  the declared ansatz.
\item \emph{Consistency check}: a nontrivial compatibility test, such as a
  monodromy product, Picard-Fuchs equation, or instanton coefficient match.
\item \emph{Heuristic/status boundary}: a physical identification not derived
  from a complete regulator construction in this monograph.
\end{enumerate}
\end{definition}

For the pure \(\mathfrak{su}(2)\) theory below, the input that the Coulomb
branch exists and is exactly described at generic points by the displayed
two-derivative Abelian Wilsonian action is
Conjecture~\ref{conj:pure-su2-sw-low-energy-solution}.  The one-loop logarithm and the
resulting monodromy at infinity are holomorphy-derived after the perturbative
Wilsonian scheme is fixed.  The residual \(u\mapsto -u\) action is
symmetry-derived.  The Picard-Lefschetz transformation at a nodal fiber is a
special-geometry theorem.  The curve
\(y^2=(x^2-\Lambda^4)(x-u)\) is constraint-derived inside the minimal
genus-one ansatz.  The hypergeometric periods, Picard-Fuchs equation,
finite-monodromy product, BPS mass vanishings, and Nekrasov coefficient are
consistency checks.  The exact nonperturbative identification of these data
with the low-energy dynamics of the four-dimensional continuum QFT is the
central conjecture; the curve algebra supplies part of its evidence ledger.

In rank one the local electromagnetic particle lattice used for central charges
is
\[
  \Gamma\simeq\mathbb Z^2,\qquad
  \gamma=(n_m,n_e),
\]
with central charge \(Z_\gamma=n_m a_D+n_e a\).  This is the lattice
\(\Gamma_{\rm part}\) of
Definition~\ref{def:pure-su2-sw-global-line-datum}: \(n_e=1\) is the electric
charge of the adjoint \(W\)-boson in the root normalization, and \(n_m=1\) is
the smooth monopole charge that becomes light at the monopole point.  The
choice of genuine Wilson--'t Hooft lines is not fixed by the Lie algebra
\(\mathfrak{su}(2)\); it is the additional datum
\(\mathcal L_\ell\).  Thus the same local period curve can be paired with the
\(SU(2)\), \(SO(3)_+\), or \(SO(3)_-\) absolute line lattice, with the theta
periodicity stated in
Definition~\ref{def:pure-su2-sw-global-line-datum}.  For two charges
\(\delta=(m_m,m_e)\) and \(\gamma=(n_m,n_e)\), use the antisymmetric pairing
\begin{equation}
  \langle \delta,\gamma\rangle
  =
  m_m n_e-m_e n_m .
  \label{eq:rank-one-em-pairing}
\end{equation}
This is the Dirac pairing in the period basis \(\Pi=(a_D,a)^T\).

\paragraph{The local coefficient in a hypermultiplet monodromy.}
The phrase ``Picard--Lefschetz monodromy of a massless hypermultiplet''
contains two logically separate inputs.  The geometric input is the
Picard--Lefschetz transvection for the vanishing one-cycle of the period
family.  The physical normalization of that transvection is fixed by the
singular Wilsonian threshold of the light charged hypermultiplet.  Let
\(\gamma\in\Gamma\) be primitive and choose an auxiliary charge
\(\eta\in\Gamma\) with \(\langle\eta,\gamma\rangle=1\).  In the local
electric frame
\[
  z=Z_\gamma,\qquad z_D=Z_\eta ,
\]
the additional light field is an ordinary \(\mathcal N=2\) hypermultiplet of
electric charge one.  Its singular holomorphic coupling is
\begin{equation}
  \tau_{\rm sing}(z)
  =
  \frac{\partial z_D}{\partial z}
  =
  -\frac{\ii}{\pi}\log z
  + \hbox{single-valued holomorphic terms}.
  \label{eq:local-hyper-singular-coupling}
\end{equation}
Equivalently,
\[
  z_D^{\rm sing}
  =
  -\frac{\ii}{\pi}z\log z
  + \hbox{single-valued holomorphic linear terms}.
\]
The coefficient in \eqref{eq:local-hyper-singular-coupling} is the
one-loop Wilsonian threshold of the two oppositely charged chiral multiplets
inside a full hypermultiplet, in the Abelian normalization of this volume.
A positive small loop around \(z=0\) therefore leaves \(z\) fixed and sends
\begin{equation}
  z_D\longmapsto z_D+2z .
  \label{eq:local-hyper-dual-period-shift}
\end{equation}
Reversing the loop orientation reverses the sign.  The factor two is not a
decorative convention: it is the local threshold coefficient of a full
hypermultiplet and is the same factor that appears in the finite
Seiberg--Witten monodromies below.

\begin{lemma}[Rank-one hypermultiplet monodromy]
\label{lem:rank-one-hyper-pl-monodromy}
Suppose that a primitive charge \(\gamma=(n_m,n_e)\) is represented by a
single massless hypermultiplet at a nodal singular fiber, and suppose the
period vector is ordered as \(\Pi=(a_D,a)^T\).  Then the Picard-Lefschetz
monodromy acting on periods is
\begin{equation}
  M_\gamma
  =
  \begin{pmatrix}
  1+2n_m n_e&2n_e^2\\
  -2n_m^2&1-2n_m n_e
  \end{pmatrix}.
  \label{eq:sw-picard-lefschetz-charge-monodromy}
\end{equation}
Equivalently, for every test charge \(\delta\),
\begin{equation}
  Z_\delta\longmapsto
  Z_\delta+2\langle\delta,\gamma\rangle Z_\gamma .
  \label{eq:rank-one-pl-central-charge-action}
\end{equation}
The matrix \(M_\gamma\) is integral symplectic and unipotent.
\end{lemma}

\begin{proof}
Choose \(\eta\in\Gamma\) with \(\langle\eta,\gamma\rangle=1\).  Every
\(\delta\in\Gamma\) can be written uniquely as
\[
  \delta=k\eta+\ell\gamma,\qquad k=\langle\delta,\gamma\rangle ,
\]
because \((\eta,\gamma)\) is an integral symplectic basis of the rank-two
lattice.  The local hypermultiplet calculation
\eqref{eq:local-hyper-dual-period-shift} gives
\[
  Z_\gamma\longmapsto Z_\gamma,\qquad
  Z_\eta\longmapsto Z_\eta+2Z_\gamma .
\]
Therefore
\[
  Z_\delta=kZ_\eta+\ell Z_\gamma
  \longmapsto
  k(Z_\eta+2Z_\gamma)+\ell Z_\gamma
  =
  Z_\delta+2\langle\delta,\gamma\rangle Z_\gamma ,
\]
which proves \eqref{eq:rank-one-pl-central-charge-action}.  The geometric
Picard--Lefschetz theorem supplies the existence of the transvection around
the vanishing cycle; the preceding Wilsonian threshold fixes its coefficient
in the physical electromagnetic lattice.

Now write \(\gamma=(p,q)\) in the original basis and apply the central-charge
action to the two basis charges.  For \(\delta=(1,0)\),
\(\langle\delta,\gamma\rangle=q\), hence
\[
  a_D'\equiv Z_{(1,0)}'
  =
  a_D+2q(pa_D+qa)
  =
  (1+2pq)a_D+2q^2a .
\]
For \(\delta=(0,1)\), \(\langle\delta,\gamma\rangle=-p\), hence
\[
  a'\equiv Z_{(0,1)}'
  =
  a-2p(pa_D+qa)
  =
  -2p^2a_D+(1-2pq)a .
\]
These two equations are exactly \eqref{eq:sw-picard-lefschetz-charge-monodromy}.
The same central-charge formula implies preservation of the antisymmetric
pairing:
\[
  \langle\delta_1+2\langle\delta_1,\gamma\rangle\gamma,
          \delta_2+2\langle\delta_2,\gamma\rangle\gamma\rangle
  =
  \langle\delta_1,\delta_2\rangle,
\]
because the two cross terms cancel and
\(\langle\gamma,\gamma\rangle=0\).  Equivalently,
\(M_\gamma^T J M_\gamma=J\) for
\[
  J=\begin{pmatrix}0&1\\ -1&0\end{pmatrix}.
\]
The determinant is one, and \(M_\gamma-I\) has square zero, so the monodromy
is unipotent.
\end{proof}

\section{Pure Gauge Theory With Gauge Algebra su two}

For the pure theory with gauge algebra \(\mathfrak{su}(2)\), the Coulomb
coordinate is
\[
  u=\left\langle \operatorname{tr}\phi^2\right\rangle .
\]
The Weyl group identifies \(a\sim -a\), and the large-field relation is
\begin{equation}
  u=\frac12a^2+O(\Lambda^4/a^2),
  \qquad
  a(u)\sim\sqrt{2u}.
  \label{eq:sw-large-u-a-u}
\end{equation}
The perturbative Wilsonian prepotential in the pure theory has the one-loop
form
\begin{equation}
  \mathcal F_{\rm pert}(a)
  =
  \frac{\ii}{2\pi}a^2
  \log\frac{a^2}{\Lambda^2}
  +
  \hbox{terms polynomial in }a .
  \label{eq:sw-perturbative-prepotential}
\end{equation}
The polynomial terms change \(a_D\) by an integral affine electric
redefinition and do not change the monodromy class.  Thus
\begin{equation}
  a_D=\frac{\partial\mathcal F}{\partial a}
  =
  \frac{\ii}{\pi}a
  \log\frac{a^2}{\Lambda^2}
  +O(a).
  \label{eq:sw-large-u-ad}
\end{equation}
When \(u\) circles infinity once counterclockwise, the square root in
\eqref{eq:sw-large-u-a-u} gives \(a\mapsto -a\), while the logarithm in
\eqref{eq:sw-large-u-ad} shifts by \(2\pi\ii\).  Therefore the period vector
\(\Pi=(a_D,a)^T\) has monodromy
\begin{equation}
  M_\infty
  =
  \begin{pmatrix}
  -1&2\\
  0&-1
  \end{pmatrix}.
  \label{eq:sw-infinity-monodromy-derived}
\end{equation}

The anomalous \(U(1)_R\) symmetry of the microscopic theory is reduced to
\(\mathbb Z_8\).  In the convention used here the vector-multiplet scalar has
\(R\)-charge \(2\), so
\[
  \phi\mapsto e^{2\ii\alpha}\phi,\qquad
  u=\langle\operatorname{tr}\phi^2\rangle
  \mapsto e^{4\ii\alpha}u .
\]
The reduction to \(\mathbb Z_8\) is the instanton-anomaly statement in these
charges.  In a one-instanton background, the index theorem gives
\(2C_2(SU(2))=4\) zero modes for each adjoint Weyl gaugino.  The
\(\mathcal N=2\) vector multiplet has two such gauginos, each of \(R\)-charge
\(1\), so the one-instanton sector acquires the phase \(\exp(8\ii\alpha)\)
under the continuous \(U(1)_R\) rotation.  The sign convention for assigning
this phase to zero-mode coordinates or to their Berezinian does not change the
invariant subgroup.  Thus only \(\alpha=\pi k/4\), \(k\in\mathbb Z\), remains
as an exact microscopic symmetry.  A generic point \(u\ne0\) is fixed by the
even values of \(k\),
namely by a \(\mathbb Z_4\) subgroup.  The quotient
\(\mathbb Z_8/\mathbb Z_4\simeq\mathbb Z_2\) therefore acts on the Coulomb
coordinate by
\[
  u\longmapsto -u .
\]
Let \(S\subset\mathbb C_u\) be the finite singular set of the rank-one period
problem.  The residual exact symmetry gives \(S=-S\).  A single finite nodal
hypermultiplet singularity is impossible in the minimal pure theory because
its Picard--Lefschetz monodromy is unipotent with trace \(2\), whereas the
weak-coupling monodromy \eqref{eq:sw-infinity-monodromy-derived} has trace
\(-2\).  The smallest symmetry-compatible nodal set away from the fixed point
\(u=0\) is therefore
\[
  S=\{s,-s\}.
\]
Since \(u\) has mass dimension \(2\), \(s\) is a nonzero dimension-two multiple
of the strong scale.  The normalization of the strong scale in this chapter is
chosen so that \(s=\Lambda^2\).  Thus the minimal two-singularity period
problem places the finite singularities at \(u=\pm\Lambda^2\).  At a nodal
singular fiber, a primitive
charge \(\gamma=(n_m,n_e)\) becomes massless and the period vector transforms
by Lemma~\ref{lem:rank-one-hyper-pl-monodromy}.  Choosing the
singularity at \(u=\Lambda^2\) to be a monopole point,
\(\gamma_m=(1,0)\), gives
\[
  M_{\Lambda^2}=M_{(1,0)}
  =
  \begin{pmatrix}
  1&0\\
  -2&1
  \end{pmatrix}.
\]
The second finite monodromy is fixed by
\[
  M_{\Lambda^2}M_{-\Lambda^2}=M_\infty,
\]
hence
\begin{equation}
  M_{-\Lambda^2}
  =
  \begin{pmatrix}
  -1&2\\
  -2&3
  \end{pmatrix}
  =
  M_{(1,-1)} .
  \label{eq:sw-dyon-monodromy-derived}
\end{equation}
Thus the second singularity carries a massless dyon of charge
\(\gamma_d=(1,-1)\) in this period basis.

\paragraph{Minimal rank-one curve inside the two-singularity ansatz.}
Assume the rank-one period problem is represented by a genus-one
hyperelliptic family with two fixed finite branch points and one branch point
depending holomorphically on \(u\).  Assume further that the only finite
singularities are exchanged by \(u\mapsto -u\), occur at
\(\pm\Lambda^2\), and are nodal singularities, while \(u=\infty\) is the
weak-coupling cusp.  Then, up to an affine change of the hyperelliptic
coordinate and a rescaling of \(\Lambda\), the minimal family is
\begin{equation}
  \Sigma_u:\qquad
  y^2=(x^2-\Lambda^4)(x-u).
  \label{eq:sw-su2-curve-derived}
\end{equation}
An affine change of \(x\) sends the two fixed finite branch points to
\(\Lambda^2\) and \(-\Lambda^2\).  A nodal singularity occurs when the moving
branch point collides with one of these fixed branch points, so the moving
branch point must take the values \(\pm\Lambda^2\) at
\(u=\pm\Lambda^2\).  In the minimal degree-one ansatz it is therefore \(x=u\)
after the same normalization of \(u\).  The fourth branch point is at
infinity, giving the weak-coupling cusp.  The resulting hyperelliptic
equation is \eqref{eq:sw-su2-curve-derived}.  Its finite branch-point
discriminant is proportional to
\[
  (\Lambda^2-(-\Lambda^2))^2
  (\Lambda^2-u)^2(-\Lambda^2-u)^2
  =
  4\Lambda^4(u^2-\Lambda^4)^2,
\]
so the only finite singular values are exactly \(u=\pm\Lambda^2\).  This
gives uniqueness only within the stated ansatz.  The QFT assertion that this
ansatz captures the exact pure theory is
Conjecture~\ref{conj:pure-su2-sw-low-energy-solution}, not a theorem of the
hyperelliptic calculation alone.

Its branch points are
\[
  x=\Lambda^2,\qquad x=-\Lambda^2,\qquad x=u,\qquad x=\infty,
\]
and the finite discriminant is proportional to
\[
  \Delta(u)=u^2-\Lambda^4 .
\]
At \(u=\Lambda^2\) the branch points \(x=u\) and \(x=\Lambda^2\) collide; at
\(u=-\Lambda^2\) the branch points \(x=u\) and \(x=-\Lambda^2\) collide.  The
monodromies of the corresponding vanishing cycles are exactly
\eqref{eq:sw-picard-lefschetz-charge-monodromy} with the charges above.  This
derivation uses the following hypotheses: the rank-one Coulomb branch has
only the two finite singularities required by the residual \(R\)-symmetry, the
singular fibers are nodal hypermultiplet singularities, and the weak-coupling
monodromy is exhausted by the perturbative vector multiplet.

\section{Periods, Picard-Fuchs Equation, and BPS Masses}

Choose the branch of the periods in the weak-coupling region
\(|u|\gg|\Lambda|^2\) by
\[
  a(u)\sim\sqrt{2u}.
\]
With the curve \eqref{eq:sw-su2-curve-derived}, the periods obey the
Picard-Fuchs equation
\begin{equation}
  (u^2-\Lambda^4)\frac{d^2\Pi}{du^2}
  +
  \frac14\Pi=0,
  \qquad
  \Pi\in\{a_D,a\}.
  \label{eq:sw-picard-fuchs-equation}
\end{equation}
The solution normalized as the electric period at infinity can be written as
\begin{equation}
  a(u)
  =
  \sqrt{2(u+\Lambda^2)}\,
  {}_2F_1
  \left(
    -\frac12,\frac12;1;
    \frac{2\Lambda^2}{u+\Lambda^2}
  \right),
  \label{eq:sw-electric-period-hypergeometric}
\end{equation}
with the branch chosen by \eqref{eq:sw-large-u-a-u}.  A dual period that
vanishes at the monopole point is
\begin{equation}
  a_D^{(m)}(u)
  =
  \frac{\ii}{2\Lambda}(u-\Lambda^2)
  {}_2F_1
  \left(
    \frac12,\frac12;2;
    \frac{\Lambda^2-u}{2\Lambda^2}
  \right).
  \label{eq:sw-monopole-period-hypergeometric}
\end{equation}
The superscript records the local branch near \(u=\Lambda^2\).  Analytic
continuation of the pair
\((a_D^{(m)},a)\) around the singularities gives the monodromy matrices above.
Near the monopole point,
\begin{equation}
  a_D^{(m)}(u)
  =
  \frac{\ii}{2\Lambda}(u-\Lambda^2)
  +O((u-\Lambda^2)^2).
  \label{eq:sw-monopole-period-linear}
\end{equation}
The monopole central charge is \(Z_{(1,0)}=a_D\), hence
\begin{equation}
  M_{(1,0)}(u)
  =
  \sqrt2\,|a_D(u)|
  =
  \frac{|u-\Lambda^2|}{\sqrt2|\Lambda|}
  +O(|u-\Lambda^2|^2).
  \label{eq:sw-monopole-mass-near-singularity}
\end{equation}
At \(u=-\Lambda^2\), use the local period basis obtained by transporting to
that singularity.  The vanishing central charge is
\[
  Z_{(1,-1)}=a_D-a,
\]
and the same nodal-fiber analysis gives
\begin{equation}
  a_D(u)-a(u)
  =
  c_-\,(u+\Lambda^2)
  +
  O((u+\Lambda^2)^2),
  \qquad
  c_-\neq0,
  \label{eq:sw-dyon-period-linear}
\end{equation}
on the dyon branch.  Therefore
\begin{equation}
  M_{(1,-1)}(u)
  =
  \sqrt2\,|a_D(u)-a(u)|
  =
  \sqrt2\,|c_-|\,|u+\Lambda^2|
  +O(|u+\Lambda^2|^2).
  \label{eq:sw-dyon-mass-near-singularity}
\end{equation}
The phase and magnitude of \(c_-\) depend on the branch and symplectic basis;
the linear vanishing and the charge \((1,-1)\) are invariant data of the
singular fiber.

\section{ADHM Moduli And The Nekrasov Expansion}
\label{sec:n2-adhm-nekrasov-expansion}

The Seiberg-Witten curve determines the low-energy prepotential indirectly
through periods.  There is also a weak-coupling construction of the same
instanton expansion from the ultraviolet gauge theory.  This section records
the construction with all convention-dependent inputs exposed.

Work in Euclidean signature on \(\mathbb R^4\simeq\mathbb C^2\), with the
orientation and instanton number convention of
Section~\ref{sec:bpst-instanton-thooft-vertex}.  The integer
\[
  k=\frac1{8\pi^2}\int_{\mathbb R^4}\operatorname{tr}(F\wedge F)
\]
is positive for the orientation chosen there.  The instanton-counting
coordinate will be denoted by \(q\).  In the common half-trace superspace
coordinate, \(q=\exp(2\pi\ii\tau_{\rm ht})\).  In the trace-delta component
coupling coordinate of Section~\ref{sec:bpst-instanton-thooft-vertex}, the
same one-instanton exponential is
\[
  q
  =
  \exp\!\left(
    -\frac{4\pi^2}{g_{\rm YM}^2}+\ii\theta
  \right)
  =
  \exp\!\left(
    -\frac{8\pi^2}{g_{\rm ht}^2}+\ii\theta
  \right).
\]
The ADHM and localization formulae below use \(q\) only as the holomorphic
coordinate that grades instanton number.  For the pure \(SU(N)\) theory,
dimensional transmutation replaces \(q\) by \(\Lambda^{2N}\) after the
renormalization scheme has been fixed.

\begin{definition}[Framed ADHM data for \(U(N)\)]
\label{def:framed-adhm-data-un}
Let \(K\simeq\mathbb C^k\) and \(W\simeq\mathbb C^N\).  Framed ADHM data are
quadruples
\[
  (B_1,B_2,I,J),
  \qquad
  B_1,B_2\in\operatorname{End}(K),\quad
  I\in\operatorname{Hom}(W,K),\quad
  J\in\operatorname{Hom}(K,W),
\]
obeying
\begin{align}
  [B_1,B_2]+IJ&=0,
  \label{eq:adhm-complex-moment-map}\\
  [B_1,B_1^\dagger]+[B_2,B_2^\dagger]+II^\dagger-J^\dagger J
  &=\zeta\,\mathbf 1_K .
  \label{eq:adhm-real-moment-map}
\end{align}
The quotient by \(U(K)\) is the hyper-K\"ahler quotient.  The parameter
\(\zeta>0\) is a stability or noncommutative resolution parameter; the
ordinary framed instanton moduli space is recovered at \(\zeta=0\) away from
small-instanton singular strata.  The framing is a trivialization at
\(S^3_\infty\).
\end{definition}

On the stable locus where the \(GL(K)\) action has no stabilizer, the framed
\(k\)-instanton ADHM moduli space for \(U(N)\) has complex dimension
\[
  2kN
\]
and real dimension \(4kN\).  Removing the center-of-mass position in
\(\mathbb R^4\) gives the centered dimension \(4kN-4\).
The complex variables \(B_1,B_2,I,J\) have complex dimension
\[
  2k^2+kN+kN=2k^2+2kN .
\]
The complex moment-map equation
\eqref{eq:adhm-complex-moment-map} is an equation in
\(\operatorname{End}(K)\), hence imposes \(k^2\) complex equations on the
smooth stable locus.  Quotienting by the complexified gauge group
\(GL(K)\) removes another \(k^2\) complex dimensions.  Thus
\[
  (2k^2+2kN)-k^2-k^2=2kN .
\]
The real moment-map quotient by \(U(K)\) is equivalent to this stable
complex quotient by the Kempf--Ness correspondence.  Therefore the real
dimension is \(2(2kN)=4kN\).  Translation of the instanton center is the
trace part of \(B_1,B_2\), a copy of \(\mathbb C^2\simeq\mathbb R^4\);
fixing it removes four real dimensions.

\begin{remark}[Gauge group and compactification]
\label{rem:adhm-u-n-su-n-compactification}
The \(U(N)\) ADHM quotient is the clean equivariant model because the framing
torus has \(N\) weights \(a_1,\ldots,a_N\).  Pure \(SU(N)\) instanton
calculations are obtained by imposing \(\sum_{\alpha=1}^Na_\alpha=0\) and by
discarding the decoupled abelian center in observables.  Ordinary
commutative \(U(1)\) gauge theory has no smooth finite-action instantons on
\(\mathbb R^4\); the \(U(1)\) fixed points in the sheaf or
noncommutative compactification are pointlike ideal-sheaf strata.  Thus the
compactification or contour prescription is part of the definition of
\(Z_{\rm inst}\), as in Definition~\ref{def:s4-nekrasov-factor}.
\end{remark}

Let
\[
  T=(\mathbb C^\times)^2_{\epsilon_1,\epsilon_2}
  \times(\mathbb C^\times)^N_{a_1,\ldots,a_N}
\]
act by rotating \(\mathbb C^2\) and by the framing torus.  The parameters
\(\epsilon_1,\epsilon_2\) are the equivariant weights of the two complex
planes, and \(a_\alpha\) are Coulomb weights.  The equivariant deformation is
called the \(\Omega\)-background.  It has two jobs: it makes the instanton
volume integral an equivariant integral, and it isolates fixed points for
localization.

\begin{quotedtheorem}[ADHM fixed points]
\label{qt:adhm-fixed-points-young-diagrams}
For the smooth framed \(U(N)\) ADHM compactification on \(\mathbb C^2\), with
generic \(a_\alpha,\epsilon_1,\epsilon_2\), the \(T\)-fixed points are
isolated and are labeled by \(N\)-tuples of Young diagrams
\[
  \vec Y=(Y_1,\ldots,Y_N),
  \qquad
  |\vec Y|:=\sum_{\alpha=1}^N |Y_\alpha|=k .
\]
The equivariant integral over the \(k\)-instanton component equals the sum of
the reciprocal equivariant Euler classes of the tangent complexes at these
fixed points.
\end{quotedtheorem}

The quoted statement is a theorem of equivariant algebraic geometry for the
framed sheaf compactification.  Its use in the physical path integral
additionally assumes that the supersymmetric regulator, the
\(\Omega\)-background, and the contour define the same equivariant integral.
This physical identification is the localization input; the dimension count
alone does not imply it.

\begin{definition}[Pure vector Nekrasov partition function]
\label{def:pure-vector-nekrasov-partition}
For pure \(U(N)\) vector multiplets, define
\[
  Z_{\rm inst}^{U(N)}(a,\epsilon_1,\epsilon_2;q)
  =
  \sum_{k\geq0}q^k Z_k^{U(N)}(a,\epsilon_1,\epsilon_2),
  \qquad
  Z_0^{U(N)}=1,
\]
where \(Z_k^{U(N)}\) is the equivariant integral over the framed
\(k\)-instanton compactification.  The one-instanton fixed-point formula is
\begin{equation}
  Z_1^{U(N)}
  =
  \frac1{\epsilon_1\epsilon_2}
  \sum_{\alpha=1}^N
  \prod_{\beta\neq\alpha}
  \frac1{(a_\alpha-a_\beta)
  (a_\alpha-a_\beta+\epsilon_1+\epsilon_2)} .
  \label{eq:nekrasov-one-instanton-un}
\end{equation}
The pure \(SU(N)\) specialization imposes \(\sum_\alpha a_\alpha=0\).  Matter
hypermultiplets multiply each fixed-point term by the equivariant Euler class
of the corresponding Dirac complex; the vector formula above is the part
needed for the pure \(SU(2)\) Seiberg-Witten comparison.
\end{definition}

\paragraph{Pure \(SU(2)\) one-instanton coefficient.}
Set \(a_1=a\), \(a_2=-a\), and
\(\epsilon=\epsilon_1+\epsilon_2\).  Then
\[
  Z_1^{SU(2)}
  =
  \frac{2}{\epsilon_1\epsilon_2(4a^2-\epsilon^2)} .
\]
Consequently, in the Nekrasov normalization
\[
  \mathcal F_{\rm inst}(a;q)
  :=
  \lim_{\epsilon_1,\epsilon_2\to0}
  \epsilon_1\epsilon_2
  \log Z_{\rm inst}^{SU(2)}(a,\epsilon_1,\epsilon_2;q),
\]
the first pure-gauge term is
\[
  \mathcal F_{\rm inst}(a;q)
  =
  \frac{q}{2a^2}+O(q^2).
\]
For the pure \(SU(2)\) theory, \(q\) is identified with \(\Lambda^4\) in the
weak-coupling scale coordinate.

Substituting \(a_1=a\), \(a_2=-a\) in
\eqref{eq:nekrasov-one-instanton-un} gives two fixed-point contributions:
\[
  Z_1^{SU(2)}
  =
  \frac1{\epsilon_1\epsilon_2}
  \left[
    \frac1{(2a)(2a+\epsilon)}
    +
    \frac1{(-2a)(-2a+\epsilon)}
  \right].
\]
The second denominator is \(2a(2a-\epsilon)\).  Hence
\[
  Z_1^{SU(2)}
  =
  \frac1{\epsilon_1\epsilon_2}
  \frac1{2a}
  \left(
    \frac1{2a+\epsilon}
    +
    \frac1{2a-\epsilon}
  \right)
  =
  \frac{2}{\epsilon_1\epsilon_2(4a^2-\epsilon^2)} .
\]
Since \(Z_{\rm inst}=1+qZ_1+O(q^2)\),
\[
  \epsilon_1\epsilon_2\log Z_{\rm inst}
  =
  q\,\frac{2}{4a^2-\epsilon^2}
  +O(q^2).
\]
Taking \(\epsilon_1,\epsilon_2\to0\) gives \(q/(2a^2)\) at order \(q\).

The coefficient above is the
first weak-coupling instanton correction to the same holomorphic prepotential
whose periods are encoded by the curve \eqref{eq:sw-su2-curve-derived},
after matching the additive quadratic ambiguity in \(\mathcal F\) and the
scale coordinate.  The comparison is stringent because the Seiberg-Witten
curve determines all powers of \(\Lambda^4/a^4\) by a period problem, while
the ADHM construction determines them by finite-dimensional equivariant
integrals.

The exponential size of the \(k\)-instanton sector is \(q^k\).  In a
transseries analysis of the weak-coupling expansion, these actions mark the
candidate locations of Borel singularities separated from the perturbative
sector by integer multiples of the one-instanton action.  A theorem-level
resurgence statement for four-dimensional supersymmetric gauge theory would
require a regulator, a summation prescription, and uniform control of the
large-order perturbative coefficients.  The assertion here is only the shared
grading convention: the same \(k\), action, and \(\theta\)-phase grade the ADHM integral,
the \(\Omega\)-background partition function, and the instanton transseries
sectors.

\section{Monodromy And Light Charges}

The weak-coupling region is the exterior of the curve of marginal stability
on which \(a_D/a\) is real.  The periods and monodromies above identify
vanishing charges and transport charges of already-known states.  They do not,
by themselves, prove existence, stability, protected degeneracy, or
completeness of a BPS spectrum.

\begin{definition}[Protected rank-one BPS index]
\label{def:rank-one-sw-bps-index}
For a chamber \(\mathcal C\subset\mathcal B^\circ\) and charge
\(\gamma\in\Gamma_{\rm part}\), the one-particle protected BPS index used here
is the second helicity supertrace
\[
  \Omega_{\mathcal C}(\gamma)
  =
  -\frac12
  \operatorname{Tr}_{\mathcal H^{\mathrm{1p,BPS}}_{\gamma,u}}
  \bigl((2J_3)^2(-1)^{2J_3}\bigr),
  \qquad u\in\mathcal C .
\]
It is locally constant inside \(\mathcal C\) and may jump only when central
charge rays align.  With this normalization an isolated hypermultiplet has
\(\Omega=+1\), and a vector multiplet has \(\Omega=-2\).
\end{definition}

\begin{conjecture}[Weak BPS-index seed]
\label{conj:sw-weak-chamber-bps-seed}
In the weak-coupling chamber of
Conjecture~\ref{conj:pure-su2-sw-low-energy-solution}, the protected
one-particle indices are
\[
  \Omega_{\rm wk}(\pm(0,1))=-2,\qquad
  \Omega_{\rm wk}(\pm(1,n))=+1\quad(n\in\mathbb Z),
\]
and vanish for the other primitive charges.  The central charges of the
displayed states are
\[
  Z_{(0,1)}=a,\qquad
  Z_{(1,n)}=a_D+na .
\]
This is the semiclassical seed expected from \(W\)-bosons and from quantizing
the monopole/dyon collective coordinates.  In this chapter it is an input, not
a consequence of the monodromy matrices.
\end{conjecture}

\begin{conjecture}[Strong BPS-index seed]
\label{conj:sw-strong-chamber-bps-seed}
Inside the strong-coupling chamber bounded by the marginal-stability curve, the
only nonzero primitive one-particle indices are
\[
  \Omega_{\rm st}(\pm(1,0))=+1,\qquad
  \Omega_{\rm st}(\pm(1,-1))=+1 .
\]
Thus the only stable primitive hypermultiplets in this chamber are the two
light states associated with the two singularities and their antiparticles.
This completeness statement is a BPS-index input unless a wall-crossing
calculation is supplied.
\end{conjecture}

The change in the BPS index can occur only when
\[
  \arg Z_{\gamma_1}=\arg Z_{\gamma_2}
\]
for charges that can form or decay as a BPS bound state.  The rank-one
Kontsevich--Soibelman wall-crossing automorphism associated to a charge
\(\gamma\) acts on formal torus coordinates by
\[
  \mathcal K_\gamma(X_\delta)
  =
  X_\delta(1-X_\gamma)^{\langle\gamma,\delta\rangle\Omega(\gamma)} .
\]
For two hypermultiplet charges \(\alpha,\beta\) with
\(\langle\alpha,\beta\rangle=1\), the elementary pentagon identity has the
form
\[
  \mathcal K_\alpha\,\mathcal K_\beta
  =
  \mathcal K_\beta\,\mathcal K_{\alpha+\beta}\,\mathcal K_\alpha ,
\]
with the product ordered by the phases of the central charges.  A theorem-level
derivation of the pure \(\mathfrak{su}(2)\) weak and strong lists would require
choosing the BPS-ray ordering and proving equality of the complete
spectrum-generator products built from the indices in
Conjectures~\ref{conj:sw-weak-chamber-bps-seed} and
\ref{conj:sw-strong-chamber-bps-seed}.  The chapter does not replace that
calculation by monodromy consistency.

\section{Argyres-Douglas Scaling As A Rank-One Singularity}
\label{sec:argyres-douglas-rank-one-singularity}

The pure \(\mathfrak{su}(2)\) theory has two separated finite singularities,
so a local duality frame exists near either one and the low-energy theory
contains a single light hypermultiplet.  A different phenomenon occurs when
two singularities with mutually nonlocal vanishing charges collide after a
mass or coupling is tuned.  If \(\gamma_1,\gamma_2\in\Gamma\) vanish at the
same point and
\[
  \langle\gamma_1,\gamma_2\rangle\neq0,
\]
then no electric-magnetic frame makes both light states purely electric.
There is therefore no local Abelian Lagrangian consisting of one vector
multiplet coupled to two ordinary electric hypermultiplets.  The
Seiberg-Witten interpretation is an interacting rank-one superconformal
fixed point, usually called an Argyres-Douglas point.  The existence of the
four-dimensional QFT fixed point is a status-boundary statement unless a
separate construction is supplied; the scaling dimensions below are
consequences of the local special-geometry model.

\begin{definition}[Rank-one cusp model]
\label{def:rank-one-ad-cusp-model}
The simplest rank-one Argyres-Douglas local period model is the cusp
\begin{equation}
  \Sigma_{u,v}:\qquad y^2=x^3+v x+u ,
  \label{eq:ad-a1-a2-local-curve}
\end{equation}
with local Seiberg-Witten differential
\begin{equation}
  \lambda_{\rm AD}=\frac{u\,dx}{y} .
  \label{eq:ad-local-differential}
\end{equation}
The parameter \(u\) is the Coulomb-branch deformation of the cusp and \(v\)
is the relevant deformation that unfolds the collision.  The normalization is
chosen so that a period of \(\lambda_{\rm AD}\) has mass dimension one.
\end{definition}

\paragraph{Discriminant of the cusp unfolding.}
For the local curve
\[
  y^2=x^3+v x+u ,
\]
the finite discriminant is
\begin{equation}
  \Delta_{\rm cusp}(u,v)=-4v^3-27u^2 .
  \label{eq:ad-cusp-discriminant}
\end{equation}
Thus the two nodal singularities in the \(u\)-plane collide at
\((u,v)=(0,0)\).  At \(v=0\), the singular fiber is a cusp rather than a
node.
For a monic cubic \(f(x)=x^3+px+q\), the discriminant is
\[
  \operatorname{disc}(f)
  =
  \prod_{i<j}(r_i-r_j)^2,
  \qquad
  f(x)=\prod_i(x-r_i),
\]
and the elementary symmetric relations give
\[
  \operatorname{disc}(x^3+px+q)=-4p^3-27q^2 .
\]
Substituting \(p=v\), \(q=u\) gives
\eqref{eq:ad-cusp-discriminant}.  For \(v\ne0\), the discriminant equation
has two local solutions for \(u\), namely two nodal degenerations after a
choice of branch of \(v^{3/2}\).  As \(v\to0\), both solutions approach
\(u=0\).  At \(u=v=0\) the polynomial is \(x^3\), so all three finite branch
points coincide and the singularity is cuspidal.

\paragraph{Mutually nonlocal light charges and electric frames.}
Let \(\Gamma\) be a rank-two electromagnetic charge lattice with symplectic
pairing \(\langle\,\cdot,\cdot\,\rangle\).  If two primitive charges
\(\gamma_1,\gamma_2\in\Gamma\) obey
\begin{equation}
  \langle\gamma_1,\gamma_2\rangle\ne0,
  \label{eq:ad-mutually-nonlocal-pairing}
\end{equation}
then no symplectic change of basis can make both charges purely electric.
Consequently a neighborhood in which both central charges vanish cannot be
described as a weakly coupled Abelian vector multiplet coupled only to
ordinary electric hypermultiplets for both charges.
In any electric-magnetic frame, the purely electric sublattice is Lagrangian:
the symplectic pairing of any two purely electric charges is zero.  A
symplectic change of basis preserves
\(\langle\,\cdot,\cdot\,\rangle\).  Therefore if both \(\gamma_1\) and
\(\gamma_2\) became purely electric in some frame, their pairing would have
to vanish in that frame and hence in every frame, contradicting
\eqref{eq:ad-mutually-nonlocal-pairing}.  The last assertion follows because
an ordinary local Abelian Lagrangian with hypermultiplets of electric charge
uses one electric polarization of \(\Gamma\); it cannot include two
hypermultiplet fields whose charges have nonzero Dirac pairing in that same
polarization.

For the local model \eqref{eq:ad-a1-a2-local-curve} with differential
\eqref{eq:ad-local-differential}, the scaling dimensions are
\begin{equation}
  [x]=\frac25,\qquad
  [y]=\frac35,\qquad
  [u]=\frac65,\qquad
  [v]=\frac45 .
  \label{eq:ad-cusp-scaling-dimensions}
\end{equation}
Thus the Coulomb-branch operator has dimension \(6/5\), and the unfolding
parameter \(v\) has dimension \(4/5\).
Quasi-homogeneity of \(y^2=x^3+v x+u\) gives
\[
  2[y]=3[x]=[u].
\]
The period of the Seiberg-Witten differential has mass dimension one, and
\(\lambda_{\rm AD}=u\,dx/y\) scales with dimension \([u]+[x]-[y]\).  Therefore
\[
  [u]+[x]-[y]=1 .
\]
Solving these equations gives \([x]=2/5\), \([y]=3/5\), and
\([u]=6/5\).  The term \(v x\) has the same dimension as \(u\), so
\([v]+[x]=6/5\), and therefore \([v]=4/5\).

This calculation is intentionally local.  It does not claim that every
collision of singular fibers produces a continuum QFT, nor does it replace
the need to specify which ultraviolet theory and deformation realize the
collision.  What it does fix is the special-geometry consequence of a cusp
with differential \(\lambda_{\rm AD}\): periods scale with dimension one,
the Coulomb operator can have fractional dimension, and mutually nonlocal
light charges force the infrared description outside an ordinary weakly
coupled Abelian hypermultiplet frame.

\section{Low-Energy Abelian Theory Near A Singular Point}
\label{sec:n2-low-energy-abelian-singular-point}

Near a point where the monopole period \(a_D\) vanishes, the correct
low-energy variables are the dual \(U(1)\) vector multiplet with scalar
\(a_D\) and the light monopole hypermultiplet.  A local Wilsonian
superpotential has the form
\[
  W_{\mathrm{loc}}=\sqrt2\,A_D M\widetilde M ,
\]
where \(A_D\) is the dual chiral multiplet and \(M,\widetilde M\) are monopole
chiral multiplets of opposite dual electric charge.  Let \(U(A_D)\) denote the
chiral coordinate whose lowest component is the Coulomb-branch coordinate
\(u\) expressed in the local dual frame.  Add a small
\(\mathcal N=1\)-preserving deformation
\begin{equation}
  W_{\mathrm{def}}=\mu\,U(A_D),
  \qquad
  \mu\in\mathbb C,\qquad |\mu|\ll |\Lambda|,
  \label{eq:sw-small-mass-deformation}
\end{equation}
so that the local superpotential is
\begin{equation}
  W_{\mathrm{loc},\mu}
  =
  \sqrt2\,A_D M\widetilde M+\mu\,U(A_D).
  \label{eq:sw-local-superpotential-with-mass}
\end{equation}
The smallness assumption keeps the analysis inside the Abelian neighborhood
where the monopole hypermultiplet is the only additional light charged
field.

\paragraph{\texorpdfstring{Monopole condensation from the local \(F\)-terms}{Monopole condensation from the local F-terms}.}
Assume \(U'(0)\neq0\) at the monopole point \(A_D=0\), and work on the local
patch described by \eqref{eq:sw-local-superpotential-with-mass}.  The
\(F\)-term equations have a branch
\begin{equation}
  A_D=0,\qquad
  M\widetilde M=-\frac{\mu}{\sqrt2}\,U'(0).
  \label{eq:sw-monopole-condensate-local}
\end{equation}
When the \(U(1)_D\) \(D\)-term equation is imposed with vanishing real
Fayet--Iliopoulos parameter,
\begin{equation}
  |M|^2-|\widetilde M|^2=0,
  \label{eq:sw-local-d-term}
\end{equation}
the dual gauge group is Higgsed for \(\mu U'(0)\neq0\).

The derivatives of \eqref{eq:sw-local-superpotential-with-mass} are
\[
  \frac{\partial W_{\mathrm{loc},\mu}}{\partial M}
  =
  \sqrt2\,A_D\widetilde M,\qquad
  \frac{\partial W_{\mathrm{loc},\mu}}{\partial \widetilde M}
  =
  \sqrt2\,A_D M,
\]
and
\[
  \frac{\partial W_{\mathrm{loc},\mu}}{\partial A_D}
  =
  \sqrt2\,M\widetilde M+\mu\,U'(A_D).
\]
The first two equations are solved by \(A_D=0\).  Substituting this value in
the third equation gives
\eqref{eq:sw-monopole-condensate-local}.  If \(U'(0)\mu\neq0\), then
\(M\widetilde M\neq0\).  Together with
\eqref{eq:sw-local-d-term}, this implies \(|M|=|\widetilde M|\neq0\), and the
oppositely charged monopole fields Higgs the dual \(U(1)_D\).

In the dual Abelian description, a Wilson line of the original electric
charge is magnetically charged with respect to \(U(1)_D\).  The Higgsed dual
gauge theory therefore supports Abrikosov--Nielsen--Olesen flux tubes whose
magnetic flux is the original electric flux.  This gives the local
confinement mechanism used in the comparison between Donaldson and
Seiberg--Witten theories.

\begin{definition}[Local SW confining-string model]
\label{def:sw-local-confining-string-datum}
In the neighborhood of a monopole singularity, a local confining-string model
consists of:
\begin{enumerate}[label=(\roman*)]
\item the dual Abelian gauge field \(A_{D,\mu}\) with curvature \(F_D\) and
  coupling \(e_D\);
\item monopole fields \(M,\widetilde M\) of dual charges \(+1\) and \(-1\);
\item the complex parameter
  \begin{equation}
    \xi_{\mathbb C}=-\frac{\mu}{\sqrt2}\,U'(0),
    \label{eq:sw-complex-fi-parameter}
  \end{equation}
  defined by the local \(F\)-term equation \(M\widetilde M=\xi_{\mathbb C}\);
\item a flux sector \(n\in\mathbb Z\) with
  \[
    \frac{1}{2\pi}\int_{\mathbb R^2}F_D=n
  \]
  on the plane transverse to the string;
\item boundary conditions at transverse infinity obeying
  \(M\widetilde M\to\xi_{\mathbb C}\) and
  \(|M|=|\widetilde M|\).
\end{enumerate}
The string tension is the infimum of the static energy per unit length in the
fixed flux sector.
\end{definition}

\begin{controlledapproximation}[BPS vortex tension in the local Abelian model]
\label{ca:sw-local-bps-vortex-tension}
If the local theory is approximated by an \(\mathcal N=2\) Abelian Higgs model
with Fayet--Iliopoulos triplet of magnitude
\(|\xi|=|\xi_{\mathbb C}|\) in the normalization
\begin{equation}
  \mathcal E
  =
  \frac{1}{2e_D^2}B_D^2+|D_1q|^2+|D_2q|^2
  +\frac{e_D^2}{2}(|q|^2-|\xi|)^2 ,
  \label{eq:sw-local-abelian-higgs-energy}
\end{equation}
then a vortex of flux \(2\pi n\) obeys
\begin{equation}
  T_n\ge 2\pi |n|\,|\xi|,
  \label{eq:sw-local-vortex-bound}
\end{equation}
and saturation is equivalent to the first-order vortex equations
\begin{equation}
  (D_1+\ii\,\operatorname{sgn}(n)D_2)q=0,
  \qquad
  B_D=\operatorname{sgn}(n)e_D^2(|\xi|-|q|^2).
  \label{eq:sw-local-vortex-equations}
\end{equation}
\end{controlledapproximation}

\paragraph{Derivation of the local vortex bound.}
For \(n>0\), complete the square in
\eqref{eq:sw-local-abelian-higgs-energy}:
\[
  \mathcal E
  =
  |(D_1+\ii D_2)q|^2
  +
  \frac{1}{2e_D^2}
  \bigl(B_D-e_D^2(|\xi|-|q|^2)\bigr)^2
  +
  |\xi|B_D
  +
  \partial_i J^i .
\]
The current \(J^i\) is a total derivative whose integral vanishes for the
finite-energy boundary conditions in
Definition~\ref{def:sw-local-confining-string-datum}.  Integrating over the
transverse plane gives
\[
  T_n\ge |\xi|\int_{\mathbb R^2}F_D=2\pi n|\xi|.
\]
For \(n<0\), use the opposite square completion.  The two square terms vanish
precisely when \eqref{eq:sw-local-vortex-equations} holds.

\section{Status Boundary}

The period construction determines an Abelian Wilsonian effective action only
after the central QFT identification in
Conjecture~\ref{conj:pure-su2-sw-low-energy-solution} is supplied.  The curve,
Picard--Fuchs equation, Picard--Lefschetz monodromies, and Nekrasov coefficient
matches are exact or finite checks of the supplied datum; they do not construct
the four-dimensional continuum theory, fix an absolute line lattice by
themselves, or prove the complete BPS-index spectrum.  The missing
nonperturbative regulator for four-dimensional supersymmetric gauge theory
remains part of the Wilsonian-scheme problem recorded in
Chapter~\ref{ch:susy-four-dimensional-n1-gauge-dynamics}.

\begin{openproblem}[Theorem-level SW construction]
\label{op:theorem-level-seiberg-witten-construction}
Construct the four-dimensional \(\mathcal N=2\) gauge theory with a regulator
and continuum limit sufficient to derive the Coulomb-branch special geometry,
period data, absolute line-lattice dependence, protected BPS indices in each
chamber, wall-crossing products, and low-energy Abelian Wilsonian action as
theorems of QFT.
\end{openproblem}
