\chapter{Quantum Spectral Curve and Hexagon Form Factors}
\label{ch:planar-n4-quantum-spectral-curve-hexagon}

\ConventionsLorentzian

The mirror TBA is exact but highly redundant.  The quantum spectral curve
compresses the same spectral information into a finite system of analytic
functions.  This chapter states the \(P\mu\)-system, explains its weak-coupling
reduction to Baxter equations, and records the separate hexagon framework for
three-point functions.

\section{Analytic T-Hook Gauge}

Before introducing the \(P\mu\)-system it is useful to isolate the analytic
T-hook step that makes the later two-row Wronskians plausible.  The input is
not the full QSC; it is the analytic Y-system data of
Chapter~\ref{ch:planar-n4-mirror-tba-y-system} together with a gauge choice for
Hirota \(T\)-functions.

\begin{assumption}[One-row analytic T-gauge]
\label{ass:qsc-one-row-t-gauge}
Use the Hirota equation
\[
  T_{a,s}^+T_{a,s}^-
  =
  T_{a+1,s}T_{a-1,s}+T_{a,s+1}T_{a,s-1},
  \qquad f^\pm(u)=f(u\pm\ii/2).
\]
In a horizontal wing of the planar \(\mathcal N=4\) T-hook, choose a gauge
with
\[
  \mathcal T_{0,m}=1,\qquad
  \mathcal T_{1,m}\in\mathcal A_m,\qquad
  \mathcal T_{1,m}(u)\to m\quad (u\to\infty),
  \qquad m\geq1,
\]
where \(\mathcal A_m=\{|\operatorname{Im}u|<m/2\}\) with the prescribed
Zhukovsky cuts removed.  Assume that the discontinuity data of the two
fermionic nodes can be represented by a real density \(\rho\) through Cauchy
transforms
\[
  G(u)=-\int\frac{\dd v}{2\pi\ii}\frac{\rho(v)}{u-v},
  \qquad \operatorname{Im}u>0,
  \qquad
  \overline G(u)=+\int\frac{\dd v}{2\pi\ii}\frac{\rho(v)}{u-v},
  \qquad \operatorname{Im}u<0.
\]
The gauge is normalized so that
\begin{equation}
  \mathcal T_{1,m}
  =
  m+G^{[m]}+\overline G^{[-m]},
  \qquad m\geq1.
  \label{eq:qsc-t-gauge-one-row}
\end{equation}
\end{assumption}

This assumption is deliberately stated as an analytic reconstruction datum.  It
summarizes the information normally hidden in the phrase ``choose the
\(\mathcal T\)-gauge'': asymptotics, analyticity strips, and the branch on which
the Cauchy transforms are evaluated.

Under Assumption~\ref{ass:qsc-one-row-t-gauge}, the Hirota equation at
\((1,m)\), \(m\geq2\), gives
\begin{equation}
  \mathcal T_{2,m}
  =
  \left(1+G^{[m+1]}-G^{[m-1]}\right)
  \left(1+\overline G^{[-m-1]}-\overline G^{[-m+1]}\right).
  \label{eq:qsc-t-gauge-two-row-factorization}
\end{equation}
Since \(\mathcal T_{0,m}=1\), Hirota gives
\[
  \mathcal T_{2,m}
  =
  \mathcal T_{1,m}^+\mathcal T_{1,m}^-
  -
  \mathcal T_{1,m+1}\mathcal T_{1,m-1}.
\]
Put
\[
  a=G^{[m+1]},\quad b=G^{[m-1]},\quad
  c=\overline G^{[-m-1]},\quad d=\overline G^{[-m+1]}.
\]
Then the right-hand side is
\[
  (m+a+d)(m+b+c)-(m+1+a+c)(m-1+b+d).
\]
Expanding and collecting terms gives
\[
  1+a-b+c-d+(a-b)(c-d)=(1+a-b)(1+c-d),
\]
which is \eqref{eq:qsc-t-gauge-two-row-factorization}.

\paragraph{Magic-sheet continuation algebra.}
Assume the one-row gauge of Assumption~\ref{ass:qsc-one-row-t-gauge}.  Suppose
that, across the mirror cut in the strip under consideration, the analytic
Y-system gluing condition
\[
  Y_{2,\pm2}(u+\ii0)=\frac{1}{Y_{1,\pm1}(u-\ii0)}
\]
imposes the Cauchy-transform boundary relation
\[
  G(u+\ii0)=-\overline G(u-\ii0).
\]
Define the magic-sheet continuation by
\[
  \widehat G=
  \begin{cases}
    G, & \text{on the branch reached from the upper half-plane},\\
    -\overline G, & \text{on the branch reached from the lower half-plane}.
  \end{cases}
\]
Then, on the continued sheet,
\begin{equation}
  \widehat{\mathcal T}_{1,m}
  =
  m+\widehat G^{[m]}-\widehat G^{[-m]},
  \qquad
  \widehat{\mathcal T}_{2,m}
  =
  \widehat{\mathcal T}_{1,1}^{[m]}
  \widehat{\mathcal T}_{1,1}^{[-m]},
  \qquad m\geq2.
  \label{eq:qsc-magic-sheet-t-gauge}
\end{equation}

The boundary relation says precisely that the upper branch \(G\) and the
negative of the lower branch \(-\overline G\) glue with the same boundary
value across the mirror cut.  Thus \(\widehat G\) is a single continued Cauchy
transform on the magic sheet; the remaining discontinuity data are carried by
the Zhukovsky short cut rather than by two independent upper/lower transforms.

In the original strip,
\[
  \mathcal T_{1,m}=m+G^{[m]}+\overline G^{[-m]} .
\]
After continuation through the mirror cut, the upper shifted term remains
\(\widehat G^{[m]}\), while the lower shifted term satisfies
\(\overline G^{[-m]}=-\widehat G^{[-m]}\).  This gives the first formula in
\eqref{eq:qsc-magic-sheet-t-gauge}.

For the two-row function use
the factorization \eqref{eq:qsc-t-gauge-two-row-factorization}.  Its first factor
becomes
\[
  1+\widehat G^{[m+1]}-\widehat G^{[m-1]}
  =
  \widehat{\mathcal T}_{1,1}^{[m]},
\]
and its second factor becomes
\[
  1+\widehat G^{[-m+1]}-\widehat G^{[-m-1]}
  =
  \widehat{\mathcal T}_{1,1}^{[-m]}.
\]
Multiplying the two factors gives the second formula.  This calculation is local in
the strip: if a pole or endpoint is crossed, the missing information is an
extra source term rather than a failure of the algebra.

The hats emphasize that these are not new functions on the same sheet but
analytic continuations through the Zhukovsky cut.  In the strip
\(|\operatorname{Im}u|<m/2\), \(\widehat{\mathcal T}_{1,m}\) agrees with the
original \(\mathcal T_{1,m}\).  Outside that strip, the equality is a statement
about a chosen continuation path.  The \(P\mu\)-system below is a finite
Riemann--Hilbert compression of this analytically continued T-hook data.

\subsection{Gauge Chain and Fermionic Discontinuities}

The stringbook derivation passes through two further analytic gauges before
introducing \(P\)-functions.  We spell out the finite algebra because the
same formulae fix the signs in the \(Y_{1,1}Y_{2,2}\) product and in the
large-\(u\) energy exponent.

\begin{assumption}[Fermionic \(\mathbf T\)-gauge and \(\mathbb T\)-gauge]
\label{ass:qsc-bold-t-and-mathbb-t-gauges}
Choose a Hirota gauge \(\mathbf T_{a,s}\) with the same Y-functions as
Chapter~\ref{ch:planar-n4-mirror-tba-y-system}.  For \(n\geq1\), its central
analyticity strips satisfy
\[
  \mathbf T_{n-1,0},\quad
  \mathbf T_{n,\pm1},\quad
  \mathbf T_{n+1,\pm2}
  \in\mathcal A_n .
\]
The boundary and left-right identifications are
\[
  \mathbf T_{n,\pm2}=\mathbf T_{2,\pm n}\quad(n\geq3),
  \qquad
  \mathbf T_{0,0}^{[+1]}=\mathbf T_{0,0}^{[-1]},
  \qquad
  \mathbf T_{0,s}=\mathbf T_{0,0}^{[s]} .
\]
On a patch where the indicated square roots are fixed, define
\begin{equation}
  \mathbb T_{a,s}
  =
  (-1)^{as}\mathbf T_{a,s}
  \left(\mathbf T_{0,0}^{[a+s]}\right)^{(a-2)/2}.
  \label{eq:qsc-mathbb-t-from-bold-t}
\end{equation}
Then \(\mathbb T_{0,s}=1\), and the magic-sheet continuations obey
\begin{align}
  \widehat{\mathbb T}_{1,m}
  &=
  \widehat h^{[m]}\widehat h^{[-m]}
  \widehat{\mathcal T}_{1,m},
  \label{eq:qsc-mathbb-t-cal-t-row-one}\\
  \widehat{\mathbb T}_{2,m}
  &=
  \widehat h^{[m+1]}\widehat h^{[m-1]}
  \widehat h^{[-m+1]}\widehat h^{[-m-1]}
  \widehat{\mathcal T}_{2,m},
  \qquad m\geq2,
  \label{eq:qsc-mathbb-t-cal-t-row-two}
\end{align}
for a function \(\widehat h\) with only the central short cut in this gauge.
\end{assumption}

The assumption records an analytic gauge choice, not a new dynamical equation.
The gauge transformation leaves the local Y-functions unchanged; its role is
to put the central row and the two horizontal wings in a form adapted to
single-short-cut Wronskians.

\paragraph{Fermionic product and central-row telescope.}
In the \(\mathbf T\)-gauge of
Assumption~\ref{ass:qsc-bold-t-and-mathbb-t-gauges},
\begin{equation}
  Y_{1,1}Y_{2,2}
  =
  \frac{\mathbf T_{1,0}}{\mathbf T_{0,0}^{[+1]}}.
  \label{eq:qsc-bold-t-fermionic-product}
\end{equation}
Moreover, for every \(n\geq1\),
\begin{equation}
  \prod_{a=1}^n
  \left(1+Y_{a,0}^{[2n-a]}\right)
  =
  \frac{
    \mathbf T_{1,0}^{[2n]}\mathbf T_{n,0}^{[n-1]}
  }{
    \mathbf T_{0,0}^{[2n-1]}\mathbf T_{n+1,0}^{[n]}
  } .
  \label{eq:qsc-central-row-telescope}
\end{equation}
Consequently the analytic-Y-system discontinuity input
\[
  \operatorname{Disc}
  \left(\log Y_{1,1}Y_{2,2}\right)^{[2n]}
  =
  \sum_{a=1}^n
  \operatorname{Disc}\log
  \left(1+Y_{a,0}^{[2n-a]}\right)
\]
is equivalently written as
\begin{equation}
  \operatorname{Disc}
  \left(\log Y_{1,1}Y_{2,2}\right)^{[2n]}
  =
  \operatorname{Disc}\log
  \frac{
    \mathbf T_{1,0}^{[2n]}\mathbf T_{n,0}^{[n-1]}
  }{
    \mathbf T_{0,0}^{[2n-1]}\mathbf T_{n+1,0}^{[n]}
  } .
  \label{eq:qsc-fermionic-discontinuity-bold-t}
\end{equation}

Use the local T-system definition
\[
  Y_{a,s}
  =
  \frac{\mathbf T_{a,s+1}\mathbf T_{a,s-1}}
       {\mathbf T_{a+1,s}\mathbf T_{a-1,s}} .
\]
Then
\[
  Y_{1,1}Y_{2,2}
  =
  \frac{\mathbf T_{1,2}\mathbf T_{1,0}}
       {\mathbf T_{2,1}\mathbf T_{0,1}}
  \frac{\mathbf T_{2,3}\mathbf T_{2,1}}
       {\mathbf T_{3,2}\mathbf T_{1,2}}
  =
  \frac{\mathbf T_{1,0}\mathbf T_{2,3}}
       {\mathbf T_{0,1}\mathbf T_{3,2}}.
\]
The \(\mathbf T\)-gauge identifications give
\(\mathbf T_{3,2}=\mathbf T_{2,3}\) and
\(\mathbf T_{0,1}=\mathbf T_{0,0}^{[+1]}\), proving
\eqref{eq:qsc-bold-t-fermionic-product}.

For the telescope, Hirota gives
\[
  1+Y_{a,0}
  =
  \frac{\mathbf T_{a,0}^+\mathbf T_{a,0}^-}
       {\mathbf T_{a+1,0}\mathbf T_{a-1,0}} .
\]
After the shift by \([2n-a]\), the \(a\)-th factor is
\[
  \frac{
    \mathbf T_{a,0}^{[2n-a+1]}
    \mathbf T_{a,0}^{[2n-a-1]}
  }{
    \mathbf T_{a+1,0}^{[2n-a]}
    \mathbf T_{a-1,0}^{[2n-a]}
  } .
\]
Multiplying from \(a=1\) to \(a=n\), every interior
\(\mathbf T_{b,0}^{[2n-b+1]}\) and
\(\mathbf T_{b,0}^{[2n-b-1]}\) with \(2\leq b\leq n-1\) appears once in the
numerator and once in the denominator.  The surviving boundary terms are
\(\mathbf T_{1,0}^{[2n]}\), \(\mathbf T_{n,0}^{[n-1]}\),
\(\mathbf T_{0,0}^{[2n-1]}\), and \(\mathbf T_{n+1,0}^{[n]}\), which is
\eqref{eq:qsc-central-row-telescope}.  Taking logarithmic discontinuities
along the central cut gives \eqref{eq:qsc-fermionic-discontinuity-bold-t}.
The first displayed discontinuity relation is the analytic-Y-system input
from the mirror TBA; the telescope above supplies its T-gauge form.

\paragraph{\(\mathbb T\)-gauge compatibility.}
Under Assumption~\ref{ass:qsc-bold-t-and-mathbb-t-gauges} and
the magic-sheet continuation algebra above,
\[
  \widehat{\mathbb T}_{2,m}
  =
  \widehat{\mathbb T}_{1,1}^{[m]}
  \widehat{\mathbb T}_{1,1}^{[-m]},
  \qquad m\geq2 .
\]
Thus the two-row Wronskian reconstruction used below may be normalized in the
\(\mathbb T\)-gauge without changing the one-row magic-sheet algebra.

Substitute
\(\widehat{\mathcal T}_{2,m}
=\widehat{\mathcal T}_{1,1}^{[m]}
\widehat{\mathcal T}_{1,1}^{[-m]}\) from
\eqref{eq:qsc-magic-sheet-t-gauge} into
\eqref{eq:qsc-mathbb-t-cal-t-row-two}.  The four \(\widehat h\)-factors
become
\[
  \widehat h^{[m+1]}\widehat h^{[m-1]}
  \widehat h^{[-m+1]}\widehat h^{[-m-1]} .
\]
Using \eqref{eq:qsc-mathbb-t-cal-t-row-one} at arguments \(u+\ii m/2\) and
\(u-\ii m/2\) gives
\[
  \widehat{\mathbb T}_{1,1}^{[m]}
  =
  \widehat h^{[m+1]}\widehat h^{[m-1]}
  \widehat{\mathcal T}_{1,1}^{[m]},
  \qquad
  \widehat{\mathbb T}_{1,1}^{[-m]}
  =
  \widehat h^{[-m+1]}\widehat h^{[-m-1]}
  \widehat{\mathcal T}_{1,1}^{[-m]} .
\]
Multiplication gives the claim.  The statement is local in the chosen
analytic gauge; zeros of \(\widehat h\) or Wronskian degeneracies are handled
by analytic continuation on the neighboring patch.

\section{Pmu System}

Let \(u\in\C\) be the spectral parameter.  The quantum spectral curve uses four
functions \(P_a(u)\), \(a=1,\ldots,4\), and an antisymmetric matrix
\(\mu_{ab}(u)=-\mu_{ba}(u)\).  The functions \(P_a\) have one short branch
cut, conventionally on \([-2g,2g]\), on their principal sheet.  Analytic
continuation through this cut is denoted by a tilde.  Introduce
\[
  P^a=\chi^{ab}P_b,
\]
where \(\chi^{ab}\) is the fixed antisymmetric tensor with nonzero entries
chosen so that \(P_aP^a=0\).

The \(P\mu\)-system is
\begin{align}
  \widetilde P_a(u)&=\mu_{ab}(u)P^b(u),
  \label{eq:planar-n4-pmu-one}\\
  \widetilde\mu_{ab}(u)-\mu_{ab}(u)
  &=
  P_a(u)\widetilde P_b(u)-P_b(u)\widetilde P_a(u),
  \label{eq:planar-n4-pmu-two}\\
  \mu_{ab}(u+\ii)&=\widetilde\mu_{ab}(u).
  \label{eq:planar-n4-pmu-period}
\end{align}
The normalization is completed by the Pfaffian condition
\begin{equation}
  \operatorname{Pf}\mu
  =
  \mu_{12}\mu_{34}-\mu_{13}\mu_{24}+\mu_{14}\mu_{23}
  =
  1,
  \label{eq:planar-n4-pmu-pfaffian}
\end{equation}
up to the standard \(P\)-basis gauge transformations.  The tilde operation is
analytic continuation through the short cut of the corresponding function;
it is not complex conjugation.  On real states one imposes an additional
reality or gluing condition relating the two sheets.

\begin{definition}[QSC Riemann--Hilbert datum]
\label{def:planar-n4-qsc-rh-datum}
Within the planar integrability framework, a \(P\mu\) Riemann--Hilbert datum
consists of the spectral plane with parameter \(u\in\mathbb C\), four
single-short-cut functions \(P_a\), an antisymmetric matrix
\(\mu_{ab}\), the fixed antisymmetric tensor \(\chi^{ab}\), and the
continuation operation \(\widetilde{\phantom P}\) through the short cut
\([-2g,2g]\).  The datum is required to satisfy
\eqref{eq:planar-n4-pmu-one}--\eqref{eq:planar-n4-pmu-period},
\eqref{eq:planar-n4-pmu-pfaffian}, regularity on the defining sheet, and
large-\(u\) asymptotics specifying the global charges.  A physical solution
adds state-dependent gluing conditions and cyclic single-trace constraints.
\end{definition}

This is a Riemann--Hilbert problem, not a finite collection of ordinary
functions on one copy of the plane.  The data above tell the reader both what
is evaluated on the defining sheet and how the functions are continued to
the neighboring sheets.

\begin{assumption}[Two-row T-hook reconstruction datum]
\label{ass:qsc-two-row-t-hook-reconstruction}
Use the shift notation
\[
  f^{[n]}(u)=f(u+n\ii/2).
\]
In the physical or magic T-hook gauge, assume that the two horizontal rows
adjacent to the central node are reconstructed from two single-short-cut
functions \(P_1,P_2\) by
\begin{equation}
  \mathbb T_{1,m}
  =
  P_1^{[m]}P_2^{[-m]}-P_2^{[m]}P_1^{[-m]},
  \qquad m\geq1,
  \label{eq:qsc-two-row-wronskian}
\end{equation}
and that Hirota continuation extends this expression to the central nodes.
Let
\[
  \mathbb T_{0,1}=\mu_{12}^2,\qquad
  \mathbb T_{3,2}=\mathbb T_{2,3}\mu_{12}.
\]
The functions \(\mathbb T_{2,1}\) and \(\mathbb T_{1,0}\) are assumed to be
regular across the central short cut in the strip where the following
calculation is made.  Finally, the Wronskian
\[
  D=P_1^{[+2]}P_2^{[-2]}-P_2^{[+2]}P_1^{[-2]}
\]
is not identically zero; any isolated zeros are handled by analytic
continuation of the resulting identities.
\end{assumption}

\paragraph{Local T-hook algebra behind the \(P\mu\) bridge.}
Under Assumption~\ref{ass:qsc-two-row-t-hook-reconstruction}, regularity of
the T-hook functions across the central short cut forces
\begin{equation}
  \widetilde\mu_{12}-\mu_{12}
  =
  P_1\widetilde P_2-P_2\widetilde P_1.
  \label{eq:qsc-local-mu12-discontinuity}
\end{equation}
Moreover
\begin{equation}
  \mathbb T_{1,0}=\mu_{12}\widetilde\mu_{12},
  \qquad
  Y_{1,1}Y_{2,2}
  =
  \frac{\widetilde\mu_{12}}{\mu_{12}}
  =
  1+
  \frac{P_1\widetilde P_2-P_2\widetilde P_1}{\mu_{12}}.
  \label{eq:qsc-local-y11-y22-bridge}
\end{equation}

The preceding assumption is the nontrivial analytic input: it asserts that a
two-row Wronskian reconstruction by single-short-cut \(P\)-functions exists
in the strip, and that the relevant central T-hook nodes are regular there.
The displayed formulae below are the local Pluecker calculation extracted
from that input.

Write \(P_a^+=P_a^{[+2]}\), \(P_a^-=P_a^{[-2]}\), and
\[
  D=P_1^+P_2^- - P_2^+P_1^- .
\]
The Hirota equation at the \((2,2)\) square, expressed through the
Wronskians in \eqref{eq:qsc-two-row-wronskian}, gives in the upper
half-plane sector
\[
  \mathbb T_{2,1}
  =
  (\widetilde P_1P_2^- - \widetilde P_2P_1^-)
  (P_1^+P_2-P_2^+P_1)
  -\mu_{12}D .
\]
Continuing across the central cut replaces \(P_1,P_2,\mu_{12}\) in the
unshifted slot by their tilded values, so the same node is
\[
  \widetilde{\mathbb T}_{2,1}
  =
  (P_1P_2^- - P_2P_1^-)
  (P_1^+\widetilde P_2-P_2^+\widetilde P_1)
  -\widetilde\mu_{12}D .
\]
Since \(\mathbb T_{2,1}\) is regular in this strip, these two expressions
are equal.  Moving the \(\mu\)-terms to one side gives
\[
  (\mu_{12}-\widetilde\mu_{12})D
  =
  A-B,
\]
where
\[
\begin{aligned}
  A&=(\widetilde P_1P_2^- - \widetilde P_2P_1^-)
    (P_1^+P_2-P_2^+P_1),\\
  B&=(P_1P_2^- - P_2P_1^-)
    (P_1^+\widetilde P_2-P_2^+\widetilde P_1).
\end{aligned}
\]
Expanding both products and canceling pairwise gives the Pluecker identity
\[
  A-B=(\widetilde P_1P_2-\widetilde P_2P_1)D.
\]
Because \(D\) is not identically zero, division by \(D\) on the open set
where it is nonzero gives
\[
  \mu_{12}-\widetilde\mu_{12}
  =
  \widetilde P_1P_2-\widetilde P_2P_1,
\]
and analyticity extends the identity through isolated zeros of \(D\).  This
is \eqref{eq:qsc-local-mu12-discontinuity}.

The \((1,1)\) Hirota square gives the central-node relation
\[
  \frac{\mathbb T_{1,0}}{\mu_{12}}
  =
  \mu_{12}+\widetilde P_2P_1-\widetilde P_1P_2 .
\]
Using the discontinuity identity just proved, the right side is
\(\widetilde\mu_{12}\), hence
\(\mathbb T_{1,0}=\mu_{12}\widetilde\mu_{12}\).  In this gauge
\[
  Y_{1,1}Y_{2,2}
  =
  \frac{\mathbb T_{1,0}}{\mathbb T_{0,1}}
  =
  \frac{\mu_{12}\widetilde\mu_{12}}{\mu_{12}^2},
\]
which proves the first equality in
\eqref{eq:qsc-local-y11-y22-bridge}.  The second equality is the same
statement after substituting
\eqref{eq:qsc-local-mu12-discontinuity}.

\paragraph{Fermionic-node asymptotics.}
Let \(x_{\rm f}(u)\) be the fermionic-node Zhukovsky sheet used in the
large-\(u\) expansion of \(Y_{1,1}Y_{2,2}\), normalized by
\[
  x_{\rm f}(u)=\frac{g}{u}+O(u^{-3}),\qquad u\to\infty .
\]
For a mirror \(Q\)-particle at rapidity \(v\), write
\[
  a=x_Q^+(v),\qquad b=x_Q^-(v),
  \qquad |a|<1<|b|,
  \qquad
  \widetilde E_Q(v)=\log\frac{b}{a},
\]
with the logarithm on the real mirror branch, and use the mirror momentum
normalization of Chapter~\ref{ch:planar-n4-mirror-tba-y-system},
\begin{equation}
  \ii\widetilde p_Q(v)=-Q-2\ii g(a-b).
  \label{eq:qsc-mirror-momentum-from-xpm}
\end{equation}
Define the fermionic-node scattering-ratio factor
\[
  R_Q(u,v)=
  \frac{x_{\rm f}(u)-a}{x_{\rm f}(u)-b}\,
  \frac{x_{\rm f}(u)^{-1}-b}{x_{\rm f}(u)^{-1}-a}.
\]
On the branch of \(\log R_Q\) which tends to
\(\log(a/b)\) at \(u=\infty\),
\begin{equation}
  \log R_Q(u,v)
  =
  -\widetilde E_Q(v)-\frac{\widetilde p_Q(v)}{u}
  +O(u^{-2}).
  \label{eq:qsc-fermionic-node-ratio-expansion}
\end{equation}
After the inverse mirror continuation of the second particle,
\(\widetilde E=-\ii p\), \(\widetilde p=-\ii E\), the same local expansion
becomes
\[
  \log R_1^{\rm phys}(u,u_j)
  =
  \ii p(u_j)+\frac{\ii E(u_j)}{u}+O(u^{-2}).
\]

Indeed, set \(x=x_{\rm f}(u)\).  Since \(x=g/u+O(u^{-3})\),
\[
  \frac{x-a}{x-b}
  =
  \frac{a}{b}
  \left(1+x\left(\frac1b-\frac1a\right)+O(x^2)\right),
  \qquad
  \frac{x^{-1}-b}{x^{-1}-a}
  =
  1+x(a-b)+O(x^2).
\]
Multiplying,
\[
  R_Q(u,v)
  =
  \frac{a}{b}
  \left[
    1+\frac{g}{u}
    \left(a-b+\frac1b-\frac1a\right)
    +O(u^{-2})
  \right].
\]
The constant term gives
\(\log(a/b)=-\widetilde E_Q\).  For the coefficient of \(u^{-1}\), use the
bound-state shortening condition
\[
  a+\frac1a-b-\frac1b=\frac{\ii Q}{g}.
\]
It implies
\[
  g\left(a-b+\frac1b-\frac1a\right)
  =
  2g(a-b)-\ii Q
  =
  -\widetilde p_Q,
\]
where the last equality is
\eqref{eq:qsc-mirror-momentum-from-xpm}.  Taking the logarithm gives
\eqref{eq:qsc-fermionic-node-ratio-expansion}.  The physical expansion is
the same identity written with the stringbook double-Wick convention
\(\widetilde E=-\ii p\), \(\widetilde p=-\ii E\).

\begin{assumption}[Fermionic-product energy normalization]
\label{ass:qsc-t-hook-charge-bridge}
Work in the analytic T-hook gauge of
Chapter~\ref{ch:planar-n4-mirror-tba-y-system}, and consider an \(SL(2)\)
single-trace state of twist \(J\), Lorentz spin \(S\), and total physical
momentum zero.  Let \(Y_{1,1}\) and \(Y_{2,2}\) denote the two fermionic
T-hook nodes adjacent to the central strip in this gauge.  The local
reconstruction algebra in
the local T-hook algebra above supplies the \(P\mu\)
expression for \(Y_{1,1}Y_{2,2}\).  The remaining framework input is the
energy normalization inherited from the excited-state mirror TBA:
\[
  \log\bigl(Y_{1,1}(u)Y_{2,2}(u)\bigr)
  =
  \frac{\ii(\Delta-J)}{u}+O(u^{-2}),
  \qquad u\to\infty
\]
in the upper-half-plane sector used for the \(SL(2)\) large-\(u\)
asymptotics.  The fermionic-node asymptotics calculation above is the
local sheet calculation behind this normalization: in the standard
excited-state mirror-TBA formula, the fermionic product is assembled from the
ratio \(R_Q\), and its \(u^{-1}\) coefficient is the same mirror-momentum
functional that appears in the energy formula
\eqref{eq:planar-n4-excited-tba-energy}.  The assumption that remains is the
global TBA contour formula for \(Y_{1,1}Y_{2,2}\), not an independent sign
choice in the QSC.
\end{assumption}

\paragraph{Dimension exponent of \(\mu_{12}\).}
Under Assumption~\ref{ass:qsc-t-hook-charge-bridge} and the \(P\mu\)
equations,
\begin{equation}
  Y_{1,1}(u)Y_{2,2}(u)
  =
  \frac{\mu_{12}(u+\ii)}{\mu_{12}(u)}.
  \label{eq:qsc-y11-y22-mu-ratio}
\end{equation}
If \(\mu_{12}\) has power-law large-\(u\) behavior on this sheet, then
\[
  \mu_{12}(u)\sim u^{\Delta-J}.
\]

The \((1,2)\) component of the \(P\mu\) discontinuity equation gives
\[
  \widetilde\mu_{12}(u)-\mu_{12}(u)
  =
  P_1(u)\widetilde P_2(u)-P_2(u)\widetilde P_1(u).
\]
Dividing by \(\mu_{12}(u)\) and using the bridge formula
\eqref{eq:qsc-local-y11-y22-bridge} gives
\[
  Y_{1,1}(u)Y_{2,2}(u)
  =
  \frac{\widetilde\mu_{12}(u)}{\mu_{12}(u)}.
\]
Pseudo-periodicity \eqref{eq:planar-n4-pmu-period} then gives
\eqref{eq:qsc-y11-y22-mu-ratio}.  Now suppose
\(\mu_{12}(u)=C u^\alpha(1+O(u^{-1}))\) with \(C\neq0\).  Then
\[
  \log\frac{\mu_{12}(u+\ii)}{\mu_{12}(u)}
  =
  \alpha\log\left(1+\frac{\ii}{u}\right)+O(u^{-2})
  =
  \frac{\ii\alpha}{u}+O(u^{-2}).
\]
Comparison with the large-\(u\) energy normalization in
Assumption~\ref{ass:qsc-t-hook-charge-bridge} gives
\(\alpha=\Delta-J\).

State quantum numbers also enter through large-\(u\) asymptotics of \(P_a\)
and the remaining \(\mu_{ab}\).  In the \(SL(2)\) sector with twist \(J\),
spin \(S\), and dimension \(\Delta\), these asymptotics encode the charges
\[
  (J,S,\Delta)
\]
through powers and leading coefficients constrained by the algebraic
relations of the QSC.
With the \(SL(2)\) conventions used in the previous chapter, the large-\(u\)
powers of the \(P\)-functions are
\[
  (P_1,P_2,P_3,P_4)
  \sim
  (A_1u^{-J/2},A_2u^{-J/2-1},A_3u^{J/2},A_4u^{J/2-1}).
\]
Together with the dimension-exponent calculation above, the leading
terms in
\[
  \widetilde P_a=\mu_{ab}P^b,\qquad
  (P^1,P^2,P^3,P^4)=(-P_4,P_3,-P_2,P_1)
\]
then force the remaining \(\mu\)-powers.  For the characteristic calculation
one should not insert the physical value of the exponent too early: keep a
trial exponent \(\alpha\) and write
\begin{equation}
  (\mu_{12},\mu_{13},\mu_{14},\mu_{23},\mu_{24},\mu_{34})
  \sim
  (u^{\alpha-J},u^{\alpha+1},u^\alpha,u^\alpha,
    u^{\alpha-1},u^{\alpha+J}),
  \label{eq:qsc-mu-leading-powers-alpha}
\end{equation}
provided the leading terms do not undergo accidental cancellations.  The
physical branch later has \(\alpha=\Delta\).

\paragraph{Large-\(u\) power balance for the \(P\mu\) sheet equation.}
Fix the \(SL(2)\) \(P\)-powers displayed above and use
\((P^1,P^2,P^3,P^4)=(-P_4,P_3,-P_2,P_1)\).  Suppose that in the trial
large-\(u\) sector \(\widetilde P_a\) has the common sheet exponent shift
\[
  \widetilde P_a=O(u^{\alpha+p_a}),
  \qquad
  (p_1,p_2,p_3,p_4)=
  \left(-\frac J2,-\frac J2-1,\frac J2,\frac J2-1\right),
\]
and that the \(\mu_{12}\) exponent is \(\alpha-J\).  If no leading row of
\(\widetilde P_a=\mu_{ab}P^b\) is cancelled identically, then the six
\(\mu\)-exponents are exactly those in
\eqref{eq:qsc-mu-leading-powers-alpha}.  With these exponents, every term
kept in the leading monodromy system
\eqref{eq:qsc-large-u-characteristic-linear-system} has the same power as
\(\mu_{ab}(u+\ii)-\mu_{ab}(u)\) in its component.

Let \(e_{ab}=e_{ba}\) be the large-\(u\) exponent of \(\mu_{ab}\), and let
\[
  (q_1,q_2,q_3,q_4)=
  \left(\frac J2-1,\frac J2,-\frac J2-1,-\frac J2\right)
\]
be the powers of \((P^1,P^2,P^3,P^4)\).  The row \(a=1\) in
\(\widetilde P_a=\mu_{ab}P^b\) has target exponent
\(\alpha-J/2\).  Since \(e_{12}=\alpha-J\),
\[
  e_{12}+q_2=\alpha-\frac J2,\qquad
  e_{13}+q_3=\alpha-\frac J2,\qquad
  e_{14}+q_4=\alpha-\frac J2,
\]
and hence \(e_{13}=\alpha+1\), \(e_{14}=\alpha\).  Applying the same balance
to the second row, whose target exponent is \(\alpha-J/2-1\), gives
\[
  e_{23}+q_3=\alpha-\frac J2-1,\qquad
  e_{24}+q_4=\alpha-\frac J2-1,
\]
so \(e_{23}=\alpha\) and \(e_{24}=\alpha-1\).  The third row has target
\(\alpha+J/2\), and therefore
\[
  e_{34}+q_4=\alpha+\frac J2,
\]
which gives \(e_{34}=\alpha+J\).  The fourth row gives the same value and is
a consistency check.  These are precisely the powers in
\eqref{eq:qsc-mu-leading-powers-alpha}.

For the monodromy recursion, the left side of the \((a,b)\) component has
power \(e_{ab}-1\), since
\[
  \mu_{ab}(u+\ii)-\mu_{ab}(u)
  =
  \ii e_{ab}m_{ab}u^{e_{ab}-1}+\cdots .
\]
Each retained term on the right is of the form
\(P_a\mu_{bc}P^c\) or \(P_b\mu_{ac}P^c\).  The row balance just proved says
\[
  e_{bc}+q_c=\alpha+p_b,\qquad
  e_{ac}+q_c=\alpha+p_a .
\]
Thus these terms have powers \(p_a+\alpha+p_b\) and \(p_b+\alpha+p_a\).
A direct substitution of the exponents above gives
\[
  p_a+p_b+\alpha=e_{ab}-1
\]
for the six pairs \((12),(13),(14),(23),(24),(34)\).  Hence the leading
linear system is homogeneous in powers of \(u\), as required.

\paragraph{Large-\(u\) characteristic determinant.}
Let
\[
  X=A_1A_4,\qquad Y=A_2A_3 .
\]
Insert the \(SL(2)\) \(P\)-powers and the trial \(\mu\)-powers
\eqref{eq:qsc-mu-leading-powers-alpha} into the one-step monodromy recursion
\eqref{eq:planar-n4-pmu-monodromy-recursion}.  For a nonzero leading vector
with \(\alpha\neq0\), the characteristic equation is
\begin{equation}
  \Phi(\alpha)
  =
  (J+1)^2\alpha^2
  -4\ii J(J+1)Y
  -
  \Bigl(
    \alpha^2+J-\ii(J+1)Y+\ii(J-1)X
  \Bigr)^2
  =
  0 .
  \label{eq:qsc-large-u-characteristic-polynomial}
\end{equation}

Write
\[
  P_1=A_1u^{-J/2},\quad
  P_2=A_2u^{-J/2-1},\quad
  P_3=A_3u^{J/2},\quad
  P_4=A_4u^{J/2-1}
\]
at leading order, and use
\((P^1,P^2,P^3,P^4)=(-P_4,P_3,-P_2,P_1)\).  If
\(\mu_{ab}=m_{ab}u^{e_{ab}}+\cdots\), then
\(\mu_{ab}(u+\ii)-\mu_{ab}(u)=\ii e_{ab}m_{ab}u^{e_{ab}-1}+\cdots\).
Keeping the leading power in each component gives the finite linear system
\begin{equation}
\begin{aligned}
  \ii(\alpha-J)m_{12}
  &=(X-Y)m_{12}+A_2^2m_{13}-A_1A_2m_{14}
    -A_1A_2m_{23}+A_1^2m_{24},\\
  \ii(\alpha+1)m_{13}
  &=-A_3^2m_{12}+(X+Y)m_{13}-A_1A_3m_{14}
    -A_1A_3m_{23}+A_1^2m_{34},\\
  \ii\alpha m_{14}
  &=-A_3A_4m_{12}+A_2A_4m_{13}
    -A_1A_3m_{24}+A_1A_2m_{34},\\
  \ii\alpha m_{23}
  &=-A_3A_4m_{12}+A_2A_4m_{13}
    -A_1A_3m_{24}+A_1A_2m_{34},\\
  \ii(\alpha-1)m_{24}
  &=-A_4^2m_{12}+A_2A_4m_{14}+A_2A_4m_{23}
    -(X+Y)m_{24}+A_2^2m_{34},\\
  \ii(\alpha+J)m_{34}
  &=-A_4^2m_{13}+A_3A_4m_{14}+A_3A_4m_{23}
    -A_3^2m_{24}+(-X+Y)m_{34}.
\end{aligned}
  \label{eq:qsc-large-u-characteristic-linear-system}
\end{equation}
Let \(\mathcal M_\alpha\) be the coefficient matrix obtained by moving the
right-hand side to the left.  Expanding the determinant and collecting only
the invariant products \(X=A_1A_4\) and \(Y=A_2A_3\) gives
\[
  \det\mathcal M_\alpha=\alpha^2\Phi(\alpha).
\]
The factor \(\alpha^2\) records a degenerate zero-exponent subspace of this
redundant six-coordinate leading system before imposing the Pfaffian and gauge
reductions.  The nonzero characteristic exponents used below therefore obey
\(\Phi(\alpha)=0\).

\begin{assumption}[Physical QSC characteristic roots]
\label{ass:qsc-large-u-characteristic-roots}
For an \(SL(2)\) single-trace state of twist \(J\), Lorentz spin \(S\), and
dimension \(\Delta\), the physical large-\(u\) gluing/asymptotic data select,
after quotienting the redundant \(\alpha^2\) factor above, the roots
\[
  \alpha=\Delta,\qquad \alpha=S-1
\]
of the characteristic polynomial
\eqref{eq:qsc-large-u-characteristic-polynomial}; cases in which the
spin-shadow root coincides with the zero sector are understood by continuity.
\end{assumption}

This root assignment is not a path-integral theorem.  It is the large-\(u\)
representation-theory and QSC-gluing input in this gauge.  The local algebra
above is included because it is a common source of sign errors:
the products \(A_1A_4\) and \(A_2A_3\) carry the sign convention for
\(\chi\), for \(f^{[\pm1]}(u)=f(u\pm\ii/2)\), and for the short-cut sheet.

\paragraph{Large-\(u\) coefficient products.}
Under Assumption~\ref{ass:qsc-large-u-characteristic-roots}, the leading
coefficient products are
\[
\begin{aligned}
  A_1A_4
  &=
  \ii\,
  \frac{((J+S-2)^2-\Delta^2)((J-S)^2-\Delta^2)}
       {16J(J-1)},\\
  A_2A_3
  &=
  \ii\,
  \frac{((J-S+2)^2-\Delta^2)((J+S)^2-\Delta^2)}
       {16J(J+1)}.
\end{aligned}
\]

Put \(\beta=S-1\).  Since \(\Phi(\Delta)=\Phi(\beta)=0\), subtracting the
two equations cancels the term proportional to \(Y\) and gives
\[
  \Delta^2+\beta^2+2J+2U=(J+1)^2,
  \qquad
  U=-\ii(J+1)Y+\ii(J-1)X,
\]
as long as \(\Delta^2\neq\beta^2\); the degenerate cases are obtained by
continuity.  Thus
\[
  U=\frac{J^2+1-\Delta^2-\beta^2}{2}.
\]
Substituting this value in \(\Phi(\Delta)=0\) gives
\[
  4\ii J(J+1)Y
  =
  (J+1)^2\Delta^2
  -
  \frac14\left(\Delta^2+(J+1)^2-\beta^2\right)^2 .
\]
Now
\[
  (J+1)^2-\beta^2=(J-S+2)(J+S),
\]
and the elementary factorization
\[
  4(J+1)^2\Delta^2
  -
  \left(\Delta^2+(J-S+2)(J+S)\right)^2
  =
  -\bigl(\Delta^2-(J-S+2)^2\bigr)
   \bigl(\Delta^2-(J+S)^2\bigr)
\]
gives the displayed formula for \(A_2A_3=Y\).  The formula for
\(A_1A_4=X\) follows from
\[
  \ii(J-1)X=U+\ii(J+1)Y
\]
and the same factorization with
\((J-S+2,J+S)\) replaced by \((J-S,J+S-2)\).

\paragraph{Consistency of the \(P\mu\) discontinuity.}
The discontinuity equation
\eqref{eq:planar-n4-pmu-two} preserves the Pfaffian normalization
\eqref{eq:planar-n4-pmu-pfaffian} if the \(P\)-functions satisfy
\[
  P_aP^a=0 .
\]

Use the rank-two Pfaffian identity
\[
  \operatorname{Pf}(M+v\wedge w)
  =
  \operatorname{Pf}(M)\bigl(1+w^TM^{-1}v\bigr)
\]
for an invertible antisymmetric \(4\times4\) matrix \(M\).  Here
\(M=\mu\), \(v=P\), and \(w=\widetilde P=\mu P^\sharp\), where
\((P^\sharp)^a=P^a=\chi^{ab}P_b\).  Hence
\[
  w^T\mu^{-1}v
  =
  (\mu P^\sharp)^T\mu^{-1}P
  =
  -(P^\sharp)^TP
  =
  -P^aP_a
  =
  0,
\]
because \(\chi^{ab}\) is antisymmetric.  Therefore crossing the cut does not
change \(\operatorname{Pf}\mu\), so the normalization can be imposed on one
sheet and transported consistently.

\paragraph{Monodromy recursion for \(\mu\).}
The \(P\mu\) equations imply the one-step continuation formula
\begin{equation}
  \mu_{ab}(u+\ii)
  =
  \mu_{ab}(u)
  +P_a(u)\mu_{bc}(u)P^c(u)
  -P_b(u)\mu_{ac}(u)P^c(u).
  \label{eq:planar-n4-pmu-monodromy-recursion}
\end{equation}
Consequently, once the functions in one strip and the continued \(P\)-sheet
are known, the shifted \(\mu\)-sheets are generated recursively by algebraic
substitution.

Use the pseudo-periodicity equation
\eqref{eq:planar-n4-pmu-period} and the discontinuity equation
\eqref{eq:planar-n4-pmu-two}:
\[
  \mu_{ab}(u+\ii)-\mu_{ab}(u)
  =
  P_a(u)\widetilde P_b(u)-P_b(u)\widetilde P_a(u).
\]
Substituting \(\widetilde P_b=\mu_{bc}P^c\) from
\eqref{eq:planar-n4-pmu-one} gives
\eqref{eq:planar-n4-pmu-monodromy-recursion}.  Applying the same identity at
successively shifted arguments gives the recursion through the tower of
short cuts.  The recursion is algebraic; it does not by itself prove that the
regularity, gluing, and large-\(u\) conditions select a unique physical
solution.

\begin{quotedtheorem}[Quantum spectral curve spectral problem]
\label{qthm:planar-n4-qsc-status}
Within the planar \(\mathcal N=4\) integrability framework, a physical
single-trace state is specified by QSC analyticity, gluing, asymptotic, and
regularity data.  Solving the resulting \(P\mu\)-system determines the exact
planar scaling dimension of the state as a function of \(\lambda\).
\end{quotedtheorem}

The theorem is quoted with its framework stated.  The QSC is a powerful exact
solution of the planar integrability system; it is not presently a
constructive definition of four-dimensional \(\mathcal N=4\) SYM.  In this
chapter, the algebraic consequences of the \(P\mu\) equations are derived
when feasible: Pfaffian preservation, monodromy recursion, the weak-coupling
Baxter degeneration, and the one-loop dimension extraction below.  The
existence of the single-short-cut \(P\)-functions, the normalization of the
discontinuity coefficient in \eqref{eq:planar-n4-pmu-two}, the physical
gluing conditions, and the equivalence to the exact planar gauge-theory
spectrum remain QSC framework inputs.

\section{Q-Omega System and Gauge Choices}

The \(P\mu\)-system is one gauge for the QSC.  A dual description uses four
functions \(Q_i(u)\), \(i=1,\ldots,4\), and an antisymmetric matrix
\(\omega_{ij}(u)\).  In the standard convention the \(Q_i\) have long cuts,
while the \(P_a\) have one short cut on the principal sheet.  The equations
mirror the \(P\mu\) system.  Introduce
\[
  Q^i=\eta^{ij}Q_j,
\]
where \(\eta^{ij}\) is a fixed antisymmetric tensor, normalized in the same
unimodular convention as the \(P\)-index tensor \(\chi^{ab}\).  Then
\[
  \widetilde Q_i=\omega_{ij}Q^j,\qquad
  \widetilde\omega_{ij}-\omega_{ij}
  =
  Q_i\widetilde Q_j-Q_j\widetilde Q_i,
  \qquad
  \omega_{ij}(u+\ii)=\widetilde\omega_{ij}(u).
\]
The dual Pfaffian normalization is
\[
  \operatorname{Pf}\omega
  =
  \omega_{12}\omega_{34}-\omega_{13}\omega_{24}
  +\omega_{14}\omega_{23}
  =
  1
\]
after the same class of \(Q\)-basis gauge transformations used below.
The two descriptions are related by the full \(Q_{A|I}\) system of
finite-difference relations.  The useful point for this monograph is not the
formal symmetry itself but the bookkeeping: choosing short cuts for \(P_a\)
and long cuts for \(Q_i\) is an analytic gauge.  Physical charges are read
from large-\(u\) asymptotics only after this gauge and the gluing conditions
are fixed.

For real left-right symmetric states, the gluing conditions may be written,
after a basis choice, as relations between \(\widetilde P_a(u)\) and complex
conjugates of \(P_a(u)\) on the real axis outside the cut.  These conditions
select the physical solution among many formal solutions of the same
finite-difference equations.  They play the same role as contour data in the
mirror TBA.

\paragraph{Dual \(Q\omega\) transport.}
Assume the local \(Q\omega\) equations displayed above and the null
contraction \(Q_iQ^i=0\).  Then, as a local algebraic consequence of these
equations, the \(\omega\)-Pfaffian is transported
consistently through a long cut, and the one-step shifted sheet obeys
\begin{equation}
  \omega_{ij}(u+\ii)
  =
  \omega_{ij}(u)
  +Q_i(u)\omega_{jk}(u)Q^k(u)
  -Q_j(u)\omega_{ik}(u)Q^k(u).
  \label{eq:qsc-qomega-monodromy-recursion}
\end{equation}

Write
\[
  \widetilde Q_i=\omega_{ij}Q^j .
\]
The discontinuity equation says that crossing the cut changes \(\omega\) by
the rank-two antisymmetric update
\[
  \omega\longmapsto \omega+Q\wedge\widetilde Q .
\]
For an invertible antisymmetric \(4\times4\) matrix \(M\),
\[
  \operatorname{Pf}(M+v\wedge w)
  =
  \operatorname{Pf}(M)\bigl(1+w^TM^{-1}v\bigr).
\]
With \(M=\omega\), \(v=Q\), and \(w=\widetilde Q=\omega Q^\sharp\),
where \((Q^\sharp)^i=Q^i=\eta^{ij}Q_j\), the correction factor is
\[
  w^T\omega^{-1}Q
  =
  (\omega Q^\sharp)^T\omega^{-1}Q
  =
  -(Q^\sharp)^TQ
  =
  -Q^iQ_i
  =
  0.
\]
Hence the Pfaffian normalization, once imposed on one sheet, is preserved by
the cut transport.  The pseudo-periodicity equation
\(\omega_{ij}(u+\ii)=\widetilde\omega_{ij}(u)\) and the discontinuity
equation give
\[
  \omega_{ij}(u+\ii)-\omega_{ij}(u)
  =
  Q_i(u)\widetilde Q_j(u)-Q_j(u)\widetilde Q_i(u).
\]
Substituting \(\widetilde Q_j=\omega_{jk}Q^k\) gives
\eqref{eq:qsc-qomega-monodromy-recursion}.

The scope of the calculation is the local transport rule after the long-cut
gauge has been chosen.  A global \(Q\omega\) Riemann--Hilbert construction
requires the cut structure, gluing data, asymptotics, and gauge fixing.  In
particular, changing the antisymmetric tensor \(\eta\), reversing the tilde
continuation, or using the opposite \(u\mapsto u-\ii\) shift convention changes
the displayed signs unless the same convention change is made everywhere in
the \(Q_{A|I}\) bridge.

\subsection{Local P-Q Bridge}

The two QSC gauges are not independent systems.  Locally in a strip where the
auxiliary \(Q\)-functions have no singularities, they are projections of a
first-order finite-difference system.  The following algebraic statement is
the part of that system that is used repeatedly when changing between
short-cut and long-cut descriptions.

\begin{assumption}[Local \(P\)-\(Q\) bridge]
\label{ass:qsc-local-pq-bridge}
Let \(Q_{a|i}(u)\), \(a,i=1,\ldots,4\), be a \(4\times4\) matrix of auxiliary
Q-functions with nonzero determinant in the strip under consideration.  Let
\(\chi^{ab}\) and \(\eta^{ij}\) be the fixed antisymmetric tensors used to
raise \(P\)- and \(Q\)-indices:
\[
  P^a=\chi^{ab}P_b,\qquad Q^i=\eta^{ij}Q_j.
\]
In the local Q-system gauge assume
\begin{align}
  Q_{a|i}^+ - Q_{a|i}^-
  &=
  P_a Q_i,
  \label{eq:qsc-local-pq-rank-one-shift}\\
  Q_i&=-P^a Q_{a|i}^-,
  \qquad
  P_a=-Q^i Q_{a|i}^-,
  \label{eq:qsc-local-pq-contractions}
\end{align}
with the null contractions
\[
  P_aP^a=0,\qquad Q_iQ^i=0.
\]
\end{assumption}

The null contractions are automatic once the upper indices are obtained from
antisymmetric forms and the same lower vector is contracted with itself.  They
are nevertheless displayed because they are exactly what makes the choice of
\(+\) or \(-\) shift harmless in the contraction formulae.

The assumption above is not a harmless notational convention.  Its
nontrivial content is the existence, in a chosen analytic strip and a chosen
local Q-system gauge, of an invertible auxiliary matrix \(Q_{a|i}\) whose
finite-difference jump has rank one and whose projections reproduce the
short-cut \(P_a\) and long-cut \(Q_i\) functions.  In the global QSC this
datum is part of the Riemann--Hilbert problem: one must specify the cut
structure, gluing conditions, large-\(u\) asymptotics, possible zeros of
\(\det Q_{a|i}\), and the residual \(\ii\)-periodic gauge transformations.
The local algebra below only records consequences of that datum after those
analytic choices have been made.

\paragraph{Unimodular \(P\)-\(Q\) bridge algebra.}
Under Assumption~\ref{ass:qsc-local-pq-bridge}, the contraction identities are
independent of the shift:
\begin{equation}
  Q_i=-P^a Q_{a|i}^+,\qquad
  P_a=-Q^i Q_{a|i}^+ .
  \label{eq:qsc-local-pq-shift-independent}
\end{equation}
Moreover
\begin{equation}
  \det(Q_{a|i}^+)=\det(Q_{a|i}^-).
  \label{eq:qsc-local-pq-determinant-preserved}
\end{equation}

Contract \eqref{eq:qsc-local-pq-rank-one-shift} with \(-P^a\).  Using
\eqref{eq:qsc-local-pq-contractions} and \(P_aP^a=0\) gives
\[
  -P^aQ_{a|i}^+
  =
  -P^aQ_{a|i}^- - P^aP_a Q_i
  =
  Q_i .
\]
Contracting the same rank-one shift with \(-Q^i\) gives
\[
  -Q^iQ_{a|i}^+
  =
  -Q^iQ_{a|i}^- - P_a Q_iQ^i
  =
  P_a ,
\]
which proves \eqref{eq:qsc-local-pq-shift-independent}.

For the determinant, write \(M_{a i}=Q_{a|i}^-\), regard \(Q^i\) as a column
vector \(Q^\sharp\), and regard \(Q_i\) as a row vector \(Q^\flat\).  The
second identity in \eqref{eq:qsc-local-pq-contractions} says
\[
  P=-M Q^\sharp .
\]
Therefore
\[
  Q_{a|i}^+
  =
  M_{a i}+P_aQ_i
  =
  \bigl(M(1-Q^\sharp Q^\flat)\bigr)_{a i}.
\]
The matrix-determinant lemma gives
\[
  \det(1-Q^\sharp Q^\flat)
  =
  1-Q^\flat Q^\sharp
  =
  1-Q_iQ^i
  =
  1.
\]
Thus \(\det(Q_{a|i}^+)=\det M=\det(Q_{a|i}^-)\).

This calculation is only local algebra.  It does not prove the existence of
global \(Q_{a|i}\) functions with the required cut structure, nor does it fix
the physical gluing conditions.  It explains why the QSC can be written in a
unimodular \(Q_{a|i}\) gauge and why the formulae
\(Q_i=-P^aQ_{a|i}^{\pm}\), \(P_a=-Q^iQ_{a|i}^{\pm}\) are conventionally quoted
without specifying the shift: the rank-one difference has zero contraction in
both null directions.  Nonconstant \(\ii\)-periodic changes of \(Q_{a|i}\)
are still analytic gauge transformations; they must be fixed before large-\(u\)
asymptotics or gluing relations are compared.

\section{Weak-Coupling Baxter Limit}

At weak coupling the short cut collapses to the origin in the rescaled
spectral plane.  In the \(SL(2)\) sector, the QSC reduces to a finite-difference
Baxter equation for a polynomial \(Q(u)\) whose zeros are one-loop Bethe roots.
The equation has the form
\begin{equation}
  (u+\ii/2)^JQ(u+\ii)
  +(u-\ii/2)^JQ(u-\ii)
  =
  T(u)Q(u),
  \label{eq:planar-n4-qsc-baxter-limit}
\end{equation}
where \(T(u)\) is a degree-\(J\) transfer polynomial fixed by charges and
regularity.  If
\[
  Q(u)=\prod_{k=1}^S(u-u_k),
\]
then evaluating \eqref{eq:planar-n4-qsc-baxter-limit} at \(u=u_j\) gives
\[
  \left(\frac{u_j+\ii/2}{u_j-\ii/2}\right)^J
  =
  \prod_{\substack{k=1\\k\neq j}}^S
  \frac{u_j-u_k-\ii}{u_j-u_k+\ii},
\]
which is the one-loop \(SL(2)\) Bethe equation in this convention.  Thus the
QSC contains the one-loop spin chain as its weak-coupling degeneration.

\begin{assumption}[Weak-coupling regularity of the \(SL(2)\) QSC]
\label{ass:qsc-weak-coupling-regularity}
Fix an \(SL(2)\) state of twist \(J\in\mathbb Z_{>0}\), Lorentz spin
\(S\in\mathbb Z_{\geq0}\), and dimension
\(\Delta=J+S+O(g^2)\).  In the weak-coupling degeneration of the
\(P\mu\)-system assume the following analytic data.
\[
  P_2P_3=O(g^2),\qquad
  P_1(u)=A_1u^{-J/2}+O(g^2),
\]
and \(P_4/P_1\) tends to a polynomial of degree \(J-1\).  Moreover
\[
  \mu_{12}(u+\ii/2)=Q(u)+g^2R(u)+O(g^4),
\]
where \(Q(u)\) is a degree-\(S\) polynomial and the singular part of \(R\) is
produced only by the short cuts \([-2g,2g]+\ii n\) collapsing to the points
\(\ii n\).  The defining sheet has no additional poles.
\end{assumption}

\paragraph{Weak \(P\mu\) elimination before the Baxter equation.}
Work in the weak-coupling regularity class of
Assumption~\ref{ass:qsc-weak-coupling-regularity}.  To leading order in
\(g\), neglect the \(P_2P_3\) terms in the \((1,2)\) and \((2,4)\)
components of the monodromy recursion
\eqref{eq:planar-n4-pmu-monodromy-recursion}.  Then
\begin{equation}
  \mu_{24}
  =
  \frac{\mu_{12}^{[+2]}-\mu_{12}}{P_1^2}
  -
  \mu_{12}\frac{P_4}{P_1},
  \label{eq:qsc-weak-mu24-from-mu12}
\end{equation}
and eliminating \(\mu_{24}\) gives
\begin{equation}
\begin{aligned}
  &\left(
    \frac{P_4^{[+1]}}{P_1^{[+1]}}
    -
    \frac{P_4^{[-1]}}{P_1^{[-1]}}
    +
    \frac{1}{(P_1^{[+1]})^2}
    +
    \frac{1}{(P_1^{[-1]})^2}
  \right)\mu_{12}^{[+1]}  \\
  &\hspace{3em}
  =
  \frac{\mu_{12}^{[+3]}}{(P_1^{[+1]})^2}
  +
  \frac{\mu_{12}^{[-1]}}{(P_1^{[-1]})^2}.
\end{aligned}
  \label{eq:qsc-weak-mu12-pre-baxter}
\end{equation}

Use \(P^1=-P_4\), \(P^4=P_1\), and discard terms proportional to \(P_2\)
or \(P_3\).  The \((1,2)\) component of
\eqref{eq:planar-n4-pmu-monodromy-recursion} gives
\[
  \mu_{12}^{[+2]}-\mu_{12}
  =
  P_1(\mu_{21}P^1+\mu_{24}P^4)
  =
  P_1(\mu_{12}P_4+\mu_{24}P_1),
\]
which is \eqref{eq:qsc-weak-mu24-from-mu12}.  The \((2,4)\) component gives
\begin{equation}
  \mu_{24}^{[+2]}-\mu_{24}
  =
  -P_4(\mu_{21}P^1+\mu_{24}P^4)
  =
  -\frac{P_4}{P_1}(\mu_{12}^{[+2]}-\mu_{12}),
  \label{eq:qsc-weak-mu24-recursion}
\end{equation}
where the last equality uses the preceding \((1,2)\) equation.  Now write
\eqref{eq:qsc-weak-mu24-from-mu12} once at \(u\) and once at \(u+\ii\), and
substitute both expressions into
\eqref{eq:qsc-weak-mu24-recursion}.  After shifting the resulting equation by
\(u\mapsto u-\ii/2\), the terms proportional to
\(\mu_{12}^{[+1]}\) collect exactly into the coefficient displayed in
\eqref{eq:qsc-weak-mu12-pre-baxter}.  This gives the stated elimination formula.

The Baxter equation obtained in this degeneration is the following.  Assume
the hypotheses of the weak \(P\mu\) elimination above.  Normalize \(P_1\) so
that, at \(g=0\),
\[
  \frac{1}{(P_1^{[\pm1]}(u))^2}=(u\pm\ii/2)^J,
\]
and assume the regular ratio \(P_4/P_1\) is such that
\[
  T(u)=
    \frac{P_4^{[+1]}}{P_1^{[+1]}}
    -
    \frac{P_4^{[-1]}}{P_1^{[-1]}}
    +(u+\ii/2)^J+(u-\ii/2)^J
\]
is a degree-\(J\) transfer polynomial.  With
\[
  Q(u)=\mu_{12}^{[+1]}(u)\big|_{g=0},
\]
the leading weak equation is the Baxter equation
\[
  (u+\ii/2)^JQ(u+\ii)
  +(u-\ii/2)^JQ(u-\ii)
  =
  T(u)Q(u).
\]
If \(Q(u)=\prod_{k=1}^S(u-u_k)\), regularity at the zeros of \(Q\) gives the
one-loop \(SL(2)\) Bethe equations.
Equation~\eqref{eq:qsc-weak-mu12-pre-baxter}, with the stated leading
asymptotics of \(P_1\), is exactly the displayed Baxter equation.  The right
side is regular at every zero \(u_j\) of \(Q\), so the left side must vanish
there:
\[
  (u_j+\ii/2)^JQ(u_j+\ii)
  +(u_j-\ii/2)^JQ(u_j-\ii)=0.
\]
Dividing by \((u_j-\ii/2)^JQ(u_j-\ii)\) and using
\[
  \frac{Q(u_j-\ii)}{Q(u_j+\ii)}
  =
  -\prod_{\substack{k=1\\k\neq j}}^S
  \frac{u_j-u_k-\ii}{u_j-u_k+\ii}
\]
gives the displayed Bethe equation.

\begin{lemma}[Half-integer digamma primitive]
\label{lem:qsc-half-integer-digamma-primitive}
Let
\[
  \Phi(u)=
  \psi\!\left(\frac12-\ii u\right)
  +\psi\!\left(\frac12+\ii u\right),
  \qquad \psi=\Gamma'/\Gamma .
\]
As a meromorphic function of \(u\), \(\Phi\) has only simple poles at
\[
  u=\pm\ii\left(n+\frac12\right),\qquad n\in\mathbb Z_{\ge0}.
\]
It satisfies the one-step shift identities
\begin{equation}
  \Phi(u+\ii)-\Phi(u)=\frac{2\ii}{u+\ii/2},
  \qquad
  \Phi(u)-\Phi(u-\ii)=\frac{2\ii}{u-\ii/2},
  \label{eq:qsc-digamma-primitive-shift-defects}
\end{equation}
and hence the central second-difference identity
\begin{equation}
  \Phi(u+\ii)-2\Phi(u)+\Phi(u-\ii)
  =
  \frac{2}{u^2+1/4}.
  \label{eq:qsc-digamma-primitive-second-difference}
\end{equation}
Moreover, if \(G\) is another meromorphic function with the same two
one-step defects in \eqref{eq:qsc-digamma-primitive-shift-defects}, then
\(G-\Phi\) is \(\ii\)-periodic wherever both functions are holomorphic.  If
this difference extends to an entire function bounded on a vertical
fundamental strip \(0\leq \operatorname{Im}u\leq1\), then \(G-\Phi\) is
constant.
\end{lemma}

\begin{proof}
The pole statement follows from the poles of \(\psi\) at
\(0,-1,-2,\ldots\).  The recurrence relation
\[
  \psi(z+1)=\psi(z)+\frac1z
\]
gives
\[
\begin{aligned}
  \Phi(u+\ii)
  &=
  \psi\!\left(\frac32-\ii u\right)
  +\psi\!\left(-\frac12+\ii u\right)  \\
  &=
  \psi\!\left(\frac12-\ii u\right)
  +\frac{1}{1/2-\ii u}
  +\psi\!\left(\frac12+\ii u\right)
  -\frac{1}{-1/2+\ii u}.
\end{aligned}
\]
Since \(1/2-\ii u=-\ii(u+\ii/2)\), this is the first identity in
\eqref{eq:qsc-digamma-primitive-shift-defects}.  For the second identity,
write
\[
\begin{aligned}
  \Phi(u-\ii)
  &=
  \psi\!\left(-\frac12-\ii u\right)
  +\psi\!\left(\frac32+\ii u\right)\\
  &=
  \psi\!\left(\frac12-\ii u\right)
  -\frac{1}{-1/2-\ii u}
  +\psi\!\left(\frac12+\ii u\right)
  +\frac{1}{1/2+\ii u},
\end{aligned}
\]
which is the second displayed defect formula after rewriting the denominators
as shifts of \(u\).  Subtracting the two one-step identities gives
\eqref{eq:qsc-digamma-primitive-second-difference}.

If \(G\) has the same one-step defects, then
\[
  (G-\Phi)(u+\ii)=(G-\Phi)(u)
\]
on the common holomorphy domain.  If this difference is entire and bounded
on a vertical fundamental strip, periodicity bounds it on all of
\(\mathbb C\); Liouville's theorem makes it constant.
\end{proof}

\paragraph{Collapsed-cut digamma package.}
\label{par:qsc-collapsed-cut-digamma-package}
Let \(Q(u)\) be a polynomial with \(Q(\ii/2)\neq0\), and set
\[
  r_0=Q'(\ii/2)-Q'(-\ii/2).
\]
Define
\begin{equation}
  R_{\rm sing}(u)
  =
  \ii r_0\,\frac{Q(u)}{Q(\ii/2)}
  \left(
    \psi\!\left(\frac12-\ii u\right)
    +\psi\!\left(\frac12+\ii u\right)
  \right),
  \qquad \psi=\Gamma'/\Gamma .
  \label{eq:qsc-digamma-singular-package}
\end{equation}
Then \(R_{\rm sing}\) has only simple poles at
\[
  u=\pm\ii\left(n+\frac12\right),\qquad n\in\mathbb Z_{\ge0},
\]
and their residues are
\[
\begin{aligned}
  \operatorname*{Res}_{u=\ii(n+1/2)}R_{\rm sing}
  &=
  -r_0\,\frac{Q(\ii(n+1/2))}{Q(\ii/2)},\\
  \operatorname*{Res}_{u=-\ii(n+1/2)}R_{\rm sing}
  &=
  r_0\,\frac{Q(-\ii(n+1/2))}{Q(\ii/2)} .
\end{aligned}
\]
In particular, if \(Q(\ii/2)=Q(-\ii/2)\), the residues at
\(\ii/2\) and \(-\ii/2\) are \(-r_0\) and \(r_0\).  In the large-\(u\)
sector used for the \(SL(2)\) asymptotics,
\[
  \frac{R_{\rm sing}(u)}{Q(u)}
  =
  2\ii\,\frac{r_0}{Q(\ii/2)}\log u+O(u^{-1}).
\]
Finally, for the meromorphic quotient \(F_{\rm sing}(u)=R_{\rm sing}(u)/Q(u)\),
\begin{equation}
  F_{\rm sing}(u+\ii)-F_{\rm sing}(u)
  =
  -\frac{2r_0}{Q(\ii/2)(u+\ii/2)},
  \qquad
  F_{\rm sing}(u)-F_{\rm sing}(u-\ii)
  =
  -\frac{2r_0}{Q(\ii/2)(u-\ii/2)} .
  \label{eq:qsc-digamma-singular-shift-defects}
\end{equation}

The residue formulas follow from the pole structure of \(\psi\).  The
digamma function has simple poles at \(0,-1,-2,\ldots\), each with residue
\(-1\).  The factor
\(\psi(1/2+\ii u)\) therefore has a pole at
\(u=\ii(n+1/2)\); since
\[
  \frac12+\ii u+n=\ii\bigl(u-\ii(n+1/2)\bigr),
\]
its residue as a function of \(u\) is \(-1/\ii=\ii\).  Multiplication by
the prefactor \(\ii r_0Q(u)/Q(\ii/2)\) gives the first residue formula.
The factor \(\psi(1/2-\ii u)\) gives the second formula in the same way,
because
\[
  \frac12-\ii u+n=-\ii\bigl(u+\ii(n+1/2)\bigr)
\]
and hence the residue in \(u\) is \(-1/(-\ii)=-\ii\).

Finally, Stirling's expansion gives
\[
  \psi(z)=\log z+O(z^{-1})
\]
in sectors avoiding the negative real axis.  Applying this to
\(z=1/2\pm\ii u\) in the chosen large-\(u\) sector gives
\[
  \psi\!\left(\frac12-\ii u\right)
  +\psi\!\left(\frac12+\ii u\right)
  =
  2\log u+O(u^{-1}),
\]
with the branch inherited from the QSC large-\(u\) sheet.  This proves the
logarithmic coefficient.  The quotient shift identities
\eqref{eq:qsc-digamma-singular-shift-defects} follow by multiplying
\eqref{eq:qsc-digamma-primitive-shift-defects} by \(\ii r_0/Q(\ii/2)\).

\paragraph{One-loop dimension from the weak QSC.}
\label{par:qsc-one-loop-dimension}
Assume the regular weak-coupling degeneration of
Assumption~\ref{ass:qsc-weak-coupling-regularity}, and impose the cyclic
single-trace condition
\[
  \frac{Q(\ii/2)}{Q(-\ii/2)}=1 .
\]
Then the first anomalous-dimension coefficient extracted from the QSC is
\[
  \Delta
  =
  J+S
  +2\ii g^2
  \left.
    \partial_u\log\frac{Q(u+\ii/2)}{Q(u-\ii/2)}
  \right|_{u=0}
  +O(g^4).
\]
Equivalently, if \(Q(u)=\prod_{k=1}^S(u-u_k)\), then
\[
  \Delta
  =
  J+S
  +2g^2\sum_{k=1}^S\frac1{u_k^2+1/4}
  +O(g^4).
\]

The leading \(P_2P_3=O(g^2)\) scaling decouples the
\((\mu_{12},\mu_{24})\) equations of the \(P\mu\)-system.  Eliminating
\(\mu_{24}\), and using the regularity of \(P_1\) and \(P_4/P_1\) in
Assumption~\ref{ass:qsc-weak-coupling-regularity}, gives the Baxter equation
derived above for
\(Q(u)=\mu_{12}(u+\ii/2)+O(g^2)\).  The same \(i\)-pseudoperiodicity of
\(\mu_{12}\), evaluated at the collapsed branch point, gives the cyclicity
condition \(Q(\ii/2)=Q(-\ii/2)\).

It remains to extract the coefficient of \(g^2\) in \(\Delta\).  Since
\(\mu_{12}(u)\sim u^{\Delta-J}\) and \(Q(u-\ii/2)\sim u^S\), the ratio
\(\mu_{12}(u)/Q(u-\ii/2)\) has the large-\(u\) expansion
\[
  1+(\Delta-J-S)\log u+O(g^4)+O(u^{-1}).
\]
The collapsed square-root cuts force the singular part of the \(g^2\)
correction to \(\mu_{12}(u+\ii/2)\) to be
\eqref{eq:qsc-digamma-singular-package}.  The collapsed-cut calculation above
shows that this expression has the
prescribed simple poles at the collapsed cuts, with residues fixed by the
branch-point relation and cyclic endpoint condition.  More precisely, after
division by the leading Baxter polynomial, the singular quotient has the
one-step defects \eqref{eq:qsc-digamma-singular-shift-defects}.  These
defects are the collapsed-cut remnant of the square-root decomposition of
\(\mu_{12}\); by Lemma~\ref{lem:qsc-half-integer-digamma-primitive}, they fix
the non-regular meromorphic primitive up to an entire \(\ii\)-periodic
ambiguity.  In the weak-coupling regularity class of
Assumption~\ref{ass:qsc-weak-coupling-regularity}, that ambiguity belongs to
the regular polynomial/asymptotic-power part of \(R\), and it carries no
\(\log u\) term.  Comparing the logarithmic coefficient in the collapsed-cut
calculation with the large-\(u\) expansion of \(\mu_{12}/Q\) gives
\[
  \Delta-J-S
  =
  2\ii g^2\,\frac{Q'(\ii/2)-Q'(-\ii/2)}{Q(\ii/2)}+O(g^4).
\]
Using \(Q(\ii/2)=Q(-\ii/2)\) rewrites this as the derivative formula.  Finally
\[
  \ii\left(
  \frac{Q'(\ii/2)}{Q(\ii/2)}
  -
  \frac{Q'(-\ii/2)}{Q(-\ii/2)}
  \right)
  =
  \sum_{k=1}^S\frac1{u_k^2+1/4},
\]
which proves the root-sum form.

\begin{example}[Konishi Baxter polynomial]
\label{ex:planar-n4-konishi-baxter-polynomial}
For the twist-two spin-two representative of Konishi, take
\[
  J=2,\qquad S=2,\qquad
  Q(u)=u^2-\frac1{12}.
\]
Then
\[
  (u+\ii/2)^2Q(u+\ii)+(u-\ii/2)^2Q(u-\ii)
  =
  \left(2u^2-\frac{13}{2}\right)Q(u).
\]
Thus the transfer polynomial is \(T(u)=2u^2-13/2\), and the zeros
\[
  u=\pm\frac{1}{2\sqrt3}
\]
are precisely the one-loop Konishi Bethe roots used in
Chapter~\ref{ch:planar-n4-spectral-problem-spin-chains}.  This example is a
finite calculation showing how the QSC weak-coupling limit remembers the
spin-chain normalization.
\end{example}

\begin{proposition}[Twist-two Baxter family]
\label{prop:planar-n4-twist-two-baxter-family}
For twist \(J=2\) and spin \(S\), define the terminating polynomial
\[
  Q_S(u)
  =
  {}_3F_2\!\left(
    \begin{matrix}
      -S,\ S+1,\ \frac12-\ii u\\
      1,\ 1
    \end{matrix}
    ;1
  \right),
\]
where \((a)_k\) is the rising Pochhammer symbol in the terminating
hypergeometric series.  It obeys
\[
  (u+\ii/2)^2Q_S(u+\ii)
  +(u-\ii/2)^2Q_S(u-\ii)
  =
  \left(2u^2-S(S+1)-\frac12\right)Q_S(u).
\]
Thus \(T_S(u)=2u^2-S(S+1)-1/2\).  The endpoint values and logarithmic
derivative are the finite identities
\[
  Q_S(-\ii/2)=1,\qquad
  Q_S(\ii/2)=(-1)^S,\qquad
  \ii\left(
  \frac{Q'_S(\ii/2)}{Q_S(\ii/2)}
  -
  \frac{Q'_S(-\ii/2)}{Q_S(-\ii/2)}
  \right)
  =
  4H_S,
\]
with \(H_S=\sum_{m=1}^S m^{-1}\).  Hence the physical even-spin twist-two
single-trace states have
\[
  \Delta_{J=2,S}
  =
  2+S+8g^2H_S+O(g^4).
\]
For \(S=2\), \(Q_2(u)=1/4-3u^2\), which is the same root set as the monic
Konishi polynomial \(u^2-1/12\), and the formula gives
\(\Delta_{\rm K}=4+12g^2+O(g^4)\).
\end{proposition}

\begin{proof}
Put
\[
  z=\frac12-\ii u,\qquad Q_S(u)=\widehat Q_S(z),
\]
so that \(u+\ii/2=\ii z\), \(u-\ii/2=\ii(z-1)\), and the shifts
\(u\mapsto u\pm\ii\) become \(z\mapsto z\pm1\).  The Baxter equation is
equivalent to
\[
  z^2\widehat Q_S(z+1)+(z-1)^2\widehat Q_S(z-1)
  =
  \bigl(2z^2-2z+S(S+1)+1\bigr)\widehat Q_S(z).
\]
Write the \(k\)-th summand of \(\widehat Q_S\) as
\[
  \tau_k(z)=
  \frac{(-S)_k(S+1)_k(z)_k}{(k!)^3},
  \qquad 0\leq k\leq S,
\]
and set \(\tau_k=0\) outside this range.  For \(k\geq1\),
\[
  (z+1)_k=(z)_k\frac{z+k}{z},
  \qquad
  (z-1)_k=(z)_k\frac{z-1}{z+k-1}.
\]
A direct substitution gives the summand identity
\[
\begin{aligned}
&z^2\tau_k(z+1)+(z-1)^2\tau_k(z-1)
  -\bigl(2z^2-2z+S(S+1)+1\bigr)\tau_k(z)  \\
&\hspace{2cm}
  =
  \frac{(k+1)^3}{z+k}\tau_{k+1}(z)
  -\frac{k^3}{z+k-1}\tau_k(z).
\end{aligned}
\]
The identity is first read away from the apparent poles \(z=1-k\) and then
extended across the removable singularities by polynomial continuation.  It is
a finite telescoping certificate.  The right side sums to zero because the
lower boundary has the factor \(k^3\) and the upper boundary has
\(\tau_{S+1}=0\).  Hence the \(z\)-form of the Baxter equation holds, and the
displayed Baxter equation follows after returning to \(u\).

The endpoint \(u=-\ii/2\) corresponds to \(z=0\), so
\(\widehat Q_S(0)=1\).  The endpoint \(u=\ii/2\) corresponds to \(z=1\), and
Chu--Vandermonde gives the finite sum
\[
  \widehat Q_S(1)
  =
  {}_2F_1(-S,S+1;1;1)
  =
  \frac{(-S)_S}{S!}=(-1)^S.
\]
It remains to prove the logarithmic-derivative identity.  Since
\[
  \frac{d}{du}Q_S(u)=-\ii\,\widehat Q'_S(z),
\]
we must compute the endpoint logarithmic derivatives of \(\widehat Q_S\).
First,
\begin{equation}
  \widehat Q'_S(0)
  =
  \sum_{k=1}^{S}
  \frac{(-S)_k(S+1)_k}{(k!)^2\,k}.
  \label{eq:qsc-twist-two-derivative-zero-sum}
\end{equation}
Indeed \(d(z)_k/dz|_{z=0}=(k-1)!\).  The same sum may be written as
\[
  \widehat Q'_S(0)
  =
  \int_0^1
  \frac{{}_2F_1(-S,S+1;1;t)-1}{t}\,dt .
\]
Using \({}_2F_1(-S,S+1;1;t)=\mathsf P_S(1-2t)\), where
\(\mathsf P_S\) is the Legendre polynomial, and setting \(x=1-2t\), this is
\[
  I_S=\int_{-1}^{1}\frac{\mathsf P_S(x)-1}{1-x}\,dx .
\]
The finite Christoffel--Darboux identity at one endpoint,
\[
  n\,\frac{\mathsf P_n(x)-\mathsf P_{n-1}(x)}{x-1}
  =
  \sum_{\ell=0}^{n-1}(2\ell+1)\mathsf P_\ell(x),
\]
is the telescoping form of the Legendre three-term recurrence.  It gives
\(I_n-I_{n-1}=-2/n\) after integration, because only the \(\ell=0\) Legendre
polynomial has nonzero integral over \([-1,1]\).  Since \(I_0=0\),
\[
  \widehat Q'_S(0)=-2H_S.
\]
At \(z=1\),
\[
  \widehat Q'_S(1)
  =
  \sum_{k=1}^{S}
  \frac{(-S)_k(S+1)_k}{(k!)^2}H_k .
\]
This is obtained from \(d(1)_k/dz=k!H_k\).  Differentiate
Chu--Vandermonde in the denominator parameter:
\[
  F(a)={}_2F_1(-S,S+1;a;1)
  =
  \frac{(a-S-1)_S}{(a)_S}.
\]
At \(a=1\),
\[
  \frac{F'(1)}{F(1)}
  =
  \sum_{m=0}^{S-1}\frac{1}{1-S-1+m}
  -
  \sum_{m=0}^{S-1}\frac{1}{1+m}
  =
  -2H_S .
\]
On the other hand differentiating the series in \(a\) gives
\[
  F'(1)=-\widehat Q'_S(1).
\]
Since \(F(1)=(-1)^S\), we obtain
\[
  \frac{\widehat Q'_S(1)}{\widehat Q_S(1)}=2H_S.
\]
Therefore
\[
  \ii\left(
  \frac{Q'_S(\ii/2)}{Q_S(\ii/2)}
  -
  \frac{Q'_S(-\ii/2)}{Q_S(-\ii/2)}
  \right)
  =
  \frac{\widehat Q'_S(1)}{\widehat Q_S(1)}
  -\frac{\widehat Q'_S(0)}{\widehat Q_S(0)}
  =
  4H_S.
\]
The dimension formula is now the specialization of
the weak-QSC one-loop dimension extraction above.
\end{proof}

\section{Cusp Anomalous Dimension and Konishi}

Two benchmark spectral quantities organize many tests of the framework.  The
first is the cusp anomalous dimension \(f(\lambda)\), extracted from large
spin twist operators:
\[
  \Delta-S
  =
  f(\lambda)\log S+O(S^0),
  \qquad S\to\infty.
\]
The QSC reproduces the same function obtained from the BES equation.  At weak
coupling it has a perturbative series in \(g^2\); at strong coupling the QSC
asymptotics begin with
\[
  f(\lambda)=\frac{\sqrt\lambda}{\pi}+O(\lambda^0)
\]
in the convention where \(\Delta-S=f(\lambda)\log S+O(S^0)\).  Comparisons
with semiclassical string theory are historically important but are not used
as inputs to the QFT-side spectral problem in this chapter.

The second is the Konishi dimension.  The QSC treats it as an ordinary finite
state with \(J=2\), \(S=2\), and zero total momentum.  At weak coupling it
matches the asymptotic Bethe ansatz until wrapping corrections enter; at
strong coupling it gives the QSC strong-coupling expansion
\[
  \Delta_{\rm K}(\lambda)
  =
  2\lambda^{1/4}
  +O(\lambda^0).
\]
The numerical and analytic solution of the QSC is a calculation problem
inside the finite functional system \eqref{eq:planar-n4-pmu-one}--\eqref{eq:planar-n4-pmu-period}.

\section{Small Spin Expansion}

The QSC gives a controlled analytic expansion near \(S=0\), where the state
approaches the protected vacuum.  The finite coefficient calculation below
begins after the following local analytic germ has been supplied.

\begin{assumption}[Small-spin QSC germ]
\label{ass:qsc-small-spin-germ}
Fix the twist \(J\in\mathbb Z_{>0}\) and let \(S\) be a complex spin parameter
near \(0\).  Assume that the physical QSC solution admits a gauge in which the
\(S=0\) limit is the protected vacuum and the first nonzero deformation has
\(\epsilon^2=O(S)\).  In this gauge the leading \(\mu\)-matrix is
\[
  \mu_{12}=-1,\quad \mu_{13}=0,\quad
  \mu_{14}=\mu_{23}=1,\quad \mu_{24}=-\sinh(2\pi u),\quad \mu_{34}=0 .
\]
Writing
\[
  x(u)=\frac{u}{2g}+\sqrt{\frac{u^2}{4g^2}-1},
\]
with the short-cut branch, the leading solution has
\[
  P_1=\epsilon x^{-J/2},
  \qquad
  P_3=\epsilon(x^{-J/2}-x^{J/2}),
\]
and the remaining two \(P\)-functions are obtained by solving the local
\(P\mu\) equations with the displayed \(\mu\)-matrix and imposing
single-short-cut regularity.  Equivalently, the Laurent expansion of
\(\sinh(2\pi u)\) is split into the powers allowed in \(P_2\) and \(P_4\).
Since
\[
  u=g(x+x^{-1}),
\]
the modified-Bessel generating function gives
\begin{equation}
  \sinh(2\pi u)
  =
  \sum_{n\in\mathbb Z} I_{2n+1}(4\pi g)x^{2n+1}.
  \label{eq:qsc-small-spin-bessel-laurent}
\end{equation}
Here \(I_m\) is the modified Bessel function of the first kind.  The first
large-\(x\) coefficients produced by the regularity split are
\[
  P_2
  =
  -\epsilon I_{J+1}(4\pi g)x^{-J/2-1}
  +O(x^{-J/2-3}),
  \qquad
  P_4
  =
  -\epsilon I_{J-1}(4\pi g)x^{J/2-1}
  +O(x^{J/2-3}).
\]
In the large-\(u\) normalization of the coefficient-product calculation above,
\[
  A_1=\epsilon g^{J/2},\quad
  A_2=-\epsilon g^{J/2+1}I_{J+1}(4\pi g),\quad
  A_3=-\epsilon g^{-J/2},\quad
  A_4=-\epsilon g^{-J/2+1}I_{J-1}(4\pi g).
\]
The corresponding large-\(u\) charge relations are
\[
  \Delta-J=-\ii(A_1A_4-A_2A_3)+O(S^2),
  \qquad
  S=-\ii(A_1A_4+A_2A_3)+O(S^2).
\]
\end{assumption}

\paragraph{Small-spin QSC slope extraction.}
\label{par:qsc-small-spin-bessel-slope}
Under Assumption~\ref{ass:qsc-small-spin-germ}, the leading anomalous
dimension satisfies
\[
  \Delta-J-S
  =
  \frac{4\pi g\,I_{J+1}(4\pi g)}{J\,I_J(4\pi g)}\,S+O(S^2).
\]

The finite calculation is as follows.  First prove the Laurent identity
\eqref{eq:qsc-small-spin-bessel-laurent}.  The generating function
\[
  \exp\left(\frac{z}{2}(t+t^{-1})\right)
  =
  \sum_{m\in\mathbb Z}I_m(z)t^m
\]
with \(z=4\pi g\) and \(t=x\) gives
\(\exp(2\pi g(x+x^{-1}))=\sum_m I_m(4\pi g)x^m\).  Replacing
\(x\) by \(-x\) gives the negative exponential.  Taking half the difference
selects exactly the odd powers and proves
\eqref{eq:qsc-small-spin-bessel-laurent}.

The displayed leading coefficients then give
\[
  A_1A_4=-\epsilon^2 g\,I_{J-1}(4\pi g),
  \qquad
  A_2A_3=\epsilon^2 g\,I_{J+1}(4\pi g).
\]
Substituting these products into the large-\(u\) charge relations in
Assumption~\ref{ass:qsc-small-spin-germ} yields
\[
  \Delta-J
  =
  \ii\epsilon^2g\bigl(I_{J-1}(4\pi g)+I_{J+1}(4\pi g)\bigr)+O(S^2),
\]
and
\[
  S
  =
  \ii\epsilon^2g\bigl(I_{J-1}(4\pi g)-I_{J+1}(4\pi g)\bigr)+O(S^2).
\]
Use the modified-Bessel recurrence
\[
  I_{J-1}(z)-I_{J+1}(z)=\frac{2J}{z}I_J(z),
  \qquad z=4\pi g.
\]
It gives
\[
  \epsilon^2
  =
  \frac{2\pi S}{\ii J I_J(4\pi g)}+O(S^2).
\]
Therefore
\[
  \Delta-J-S
  =
  2\ii\epsilon^2g\,I_{J+1}(4\pi g)+O(S^2)
  =
  \frac{4\pi g\,I_{J+1}(4\pi g)}{J\,I_J(4\pi g)}S+O(S^2),
\]
which is the stated slope.

Continuing the resulting strong-coupling expansion to \(J=2,S=2\) gives the
Konishi estimate
\[
  \Delta_{\rm K}
  =
  2\lambda^{1/4}+2\lambda^{-1/4}+O(\lambda^{-3/4}).
\]
The analytic continuation from small spin to integer spin is a QSC input, not
a theorem proved from the gauge-theory path integral.

\section{Hexagon Form Factors}

The spectral problem determines scaling dimensions.  Three-point functions
require additional data.  In the planar limit, a single-trace three-point
function can be represented by a pair of pants worldsheet.  The hexagon
program cuts this pair of pants into two hexagons and assigns a form factor to
each half.

For three single-trace operators with spin-chain lengths \(L_1,L_2,L_3\), the
bridge lengths are
\[
  \ell_{12}=\frac{L_1+L_2-L_3}{2},\qquad
  \ell_{23}=\frac{L_2+L_3-L_1}{2},\qquad
  \ell_{31}=\frac{L_3+L_1-L_2}{2}.
\]

\begin{remark}[Planar bridge-length bookkeeping]
\label{rem:planar-n4-hexagon-bridge-lengths}
Consider the planar tree-level pair-of-pants skeleton for three cyclic
single-trace operators of lengths \(L_1,L_2,L_3\).  Let \(\ell_{ij}=\ell_{ji}\)
be the number of adjacent Wick-contraction slots shared by operators \(i\)
and \(j\).  Then the bridge lengths are forced to be
\begin{equation}
  \ell_{12}=\frac{L_1+L_2-L_3}{2},\qquad
  \ell_{23}=\frac{L_2+L_3-L_1}{2},\qquad
  \ell_{31}=\frac{L_3+L_1-L_2}{2}.
  \label{eq:planar-n4-hexagon-bridge-lengths-derived}
\end{equation}
Conversely, a nonnegative integer planar skeleton of this type exists exactly
when \(L_1+L_2+L_3\) is even and the three triangle inequalities
\(L_i+L_j\geq L_k\) hold.

In the planar skeleton each trace is cut at two cyclic splitting points.  The
letters of operator \(1\) are divided into the bridge shared with operator
\(2\) and the bridge shared with operator \(3\), and similarly for the other
two operators.  Hence
\[
  L_1=\ell_{12}+\ell_{31},\qquad
  L_2=\ell_{12}+\ell_{23},\qquad
  L_3=\ell_{23}+\ell_{31}.
\]
Adding the first two equations and subtracting the third gives
\(2\ell_{12}=L_1+L_2-L_3\).  The other two formulae follow by cyclic
permutation.  Since \(\ell_{ij}\) counts letters, the right-hand sides must
be nonnegative integers; this is precisely even total length plus the
triangle inequalities.  Conversely, if these conditions hold, the three
numbers in \eqref{eq:planar-n4-hexagon-bridge-lengths-derived} define
nonnegative integer arcs on the three cyclic words satisfying the displayed
length-balance equations, which is the required pair-of-pants skeleton.
\end{remark}

Asymptotically, the structure constant is a sum over partitions of magnons
between the two hexagons:
\[
  C_{123}^{\rm asym}
  =
  \sum_{\alpha\cup\bar\alpha=\{u\}}
  (-1)^{|\bar\alpha|}
  \ee^{\ii p(\bar\alpha)\ell}
  H(\alpha)H(\bar\alpha)\,
  \mathcal N^{-1/2}.
\]
Here the displayed formula is the specialization in which the magnons shown
belong to one external operator: \(\ell\) is the bridge crossed by the subset
\(\bar\alpha\), \(H\) is the hexagon form factor, and
\[
  p(\bar\alpha)=\sum_{u_j\in\bar\alpha}p(u_j).
\]
The phase is the spin-chain translation phase.  Moving a magnon of momentum
\(p_j\) through \(\ell\) vacuum letters multiplies the coordinate Bethe wave
by \(\ee^{\ii p_j\ell}\), and moving all magnons in \(\bar\alpha\) gives
\begin{equation}
  \prod_{u_j\in\bar\alpha}\ee^{\ii p_j\ell}
  =
  \ee^{\ii p(\bar\alpha)\ell}.
  \label{eq:planar-n4-hexagon-bridge-translation-phase}
\end{equation}
The factor \(\mathcal N\) is the Gaudin norm factor of the Bethe state.
Mirror particles on the bridges add finite-size corrections, analogous in
spirit to wrapping corrections in the spectral problem.

For scalar magnons the elementary two-particle scalar factor entering the
hexagon can be written, in the same \(x^\pm\) and \(\sigma\) conventions as
Chapter~\ref{ch:planar-n4-asymptotic-bethe-ansatz}, as
\begin{equation}
  h(u,v)
  =
  \frac{x_u^- -x_v^-}{x_u^- -x_v^+}
  \frac{1-1/(x_u^+x_v^-)}{1-1/(x_u^+x_v^+)}
  \frac{1}{\sigma(u,v)} .
  \label{eq:planar-n4-hexagon-scalar-factor}
\end{equation}
The matrix part is the same centrally extended symmetry intertwiner that
appears in the magnon S-matrix, after the crossing moves needed to put
excitations on a chosen hexagon edge.  Thus the hexagon construction imports
the same analytic-continuation obligations: specifying \(h(u,v)\) without the
crossing path of the moved excitation is incomplete.

\paragraph{Hexagon scalar Watson factor.}
Let \(x_i^\pm\), \(i=1,2\), be non-zero Zhukovsky variables in a sheet domain
where the denominators below do not vanish, and write
\(\sigma_{ij}=\sigma(x_i^\pm,x_j^\pm)\).  Assume scalar dressing unitarity
\(\sigma_{12}\sigma_{21}=1\) on this domain.  If \(h_{12}=h(u_1,u_2)\) is
defined by \eqref{eq:planar-n4-hexagon-scalar-factor}, then
\begin{equation}
  \frac{h_{12}}{h_{21}}
  =
  S^{\rm hex}_{\rm scal}(1,2),
  \qquad
  S^{\rm hex}_{\rm scal}(1,2)
  =
  \frac{x_1^+ -x_2^-}{x_1^- -x_2^+}
  \frac{1-1/(x_1^+x_2^-)}{1-1/(x_1^-x_2^+)}
  \frac{1}{\sigma_{12}^2}.
  \label{eq:planar-n4-hexagon-watson-scalar}
\end{equation}
On the weak physical branch
\[
  x_i^\pm=\frac{u_i\pm\ii/2}{g}+O(g),
  \qquad
  \sigma_{12}=1+O(g^6),
\]
with \(u_1-u_2\neq\ii\), this becomes
\begin{equation}
  S^{\rm hex}_{\rm scal}(1,2)
  =
  \frac{u_1-u_2+\ii}{u_1-u_2-\ii}
  +O(g^2).
  \label{eq:planar-n4-hexagon-watson-weak-limit}
\end{equation}

From \eqref{eq:planar-n4-hexagon-scalar-factor},
\[
\begin{aligned}
  \frac{h_{12}}{h_{21}}
  &=
  \frac{x_1^- -x_2^-}{x_1^- -x_2^+}
  \frac{x_2^- -x_1^+}{x_2^- -x_1^-}
  \frac{1-1/(x_1^+x_2^-)}{1-1/(x_1^+x_2^+)}
  \frac{1-1/(x_2^+x_1^+)}{1-1/(x_2^+x_1^-)}
  \frac{\sigma_{21}}{\sigma_{12}} .
\end{aligned}
\]
The first two rational factors reduce to
\[
  \frac{x_1^- -x_2^-}{x_2^- -x_1^-}
  \frac{x_2^- -x_1^+}{x_1^- -x_2^+}
  =
  \frac{x_1^+ -x_2^-}{x_1^- -x_2^+}.
\]
The middle denominator \(1-1/(x_1^+x_2^+)\) cancels the identical factor
\(1-1/(x_2^+x_1^+)\), and the remaining denominator is
\(1-1/(x_1^-x_2^+)\).  Finally scalar dressing unitarity gives
\(\sigma_{21}/\sigma_{12}=1/\sigma_{12}^2\).  This proves
\eqref{eq:planar-n4-hexagon-watson-scalar}.

For the weak limit, the first rational factor is
\[
  \frac{(u_1+\ii/2)/g-(u_2-\ii/2)/g+O(g)}
       {(u_1-\ii/2)/g-(u_2+\ii/2)/g+O(g)}
  =
  \frac{u_1-u_2+\ii}{u_1-u_2-\ii}+O(g^2).
\]
The second rational factor is \(1+O(g^2)\), because
\((x_i^\pm x_j^\pm)^{-1}=O(g^2)\), and the dressing factor contributes
\(1+O(g^6)\).  This proves
\eqref{eq:planar-n4-hexagon-watson-weak-limit}.

The weak factor in
\eqref{eq:planar-n4-hexagon-watson-weak-limit} is the compact one-loop
exchange phase.  It should not be identified with the rank-one \(SL(2)\)
phase in \eqref{eq:planar-n4-rank-one-stringbook-aba} unless the vacuum,
edge-crossing path, and scalar-factor split have all been transported to the
same convention.

\begin{openproblem}[Derivation of the hexagon expansion]
\label{op:planar-n4-hexagon-derivation}
Give a derivation of the hexagon expansion from a nonperturbatively defined
planar gauge-theory path integral, including the emergence of the hexagon
form factor, the measure for mirror corrections, and the treatment of
operator mixing at finite length.  Existing evidence is strong inside the
integrability framework.  A derivation at the level of a theorem in
four-dimensional QFT remains open.
\end{openproblem}

\section{Maldacena--Wilson Line, Cusp, and Bremsstrahlung}

Let \(C\) be an oriented smooth curve with parameter \(s\), tangent
\(\dot x^\mu(s)\), and scalar-coupling unit vector
\(\theta^I(s)\in S^5\).  The Maldacena--Wilson line in the fundamental
representation is
\[
  W_{\rm MW}(C,\theta)
  =
  \operatorname{Tr}\,P\exp\int_C
  \bigl(\ii A_\mu(x(s))\dot x^\mu(s)
  +\Phi_I(x(s))\theta^I(s)|\dot x(s)|\bigr)\,\dd s .
\]
It is a gauge-invariant line operator after taking the trace and imposing the
usual endpoint prescription for closed curves or external heavy sources.  A
straight line with constant \(\theta\) is half-BPS and has vanishing cusp
anomalous dimension.

For two rays meeting with Euclidean deflection angle \(\varphi\), define
\(\Gamma_{\rm cusp}(\varphi,\theta)\) by the logarithmic divergence of the
renormalized cusped line:
\[
  \log\langle W_{\rm cusp}\rangle_{\rm ren}
  =
  -\Gamma_{\rm cusp}(\varphi)\log\frac{R}{\epsilon}+O(1),
\]
where \(R\) is an infrared distance scale and \(\epsilon\) is the short-distance
cutoff at the cusp.  At small geometric angle and fixed scalar coupling,
\[
  \Gamma_{\rm cusp}(\varphi)
  =
  -B(\lambda)\varphi^2+O(\varphi^4).
\]
The coefficient \(B(\lambda)\) is the Bremsstrahlung function.  On the
straight-line defect, the transverse displacement operators \(\mathbb D_i(t)\)
are defined by the first shape variation of the line; \(i=1,2,3\) labels
directions normal to the straight Euclidean line.  Conformal invariance on the
defect fixes their two-point function to
\[
  \langle \mathbb D_i(t)\mathbb D_j(0)\rangle
  =
  \frac{C_D\delta_{ij}}{(t-\ii0)^4}.
\]

\begin{proposition}[Displacement normalization of the small cusp]
\label{prop:displacement-normalization-small-cusp}
Assume the straight half-BPS line admits the standard defect shape-variation
Ward identity: for a small transverse deformation \(X^i(t)\) of compact
support, the universal nonlocal quadratic part of
\(\log\langle W_X\rangle\) is, with summation over the transverse index \(i\),
\[
  \frac12
  \int\frac{\dd\omega}{2\pi}\,
  \frac{\pi C_D}{6}|\omega|^3
  \widehat X^i(\omega)\widehat X^i(-\omega),
\]
up to local counterterms polynomial in \(\omega\).  Equivalently, this fixes
the nonlocal part of the finite-part Fourier transform of the displacement
two-point distribution,
\[
  \operatorname{Fp}\int\dd t\,
  \frac{\ee^{\ii\omega t}}{(t-\ii0)^4}
  =
  \frac{\pi}{6}|\omega|^3
  +\text{a polynomial in }\omega .
\]
Then the small-angle cusp coefficient satisfies
\[
  C_D=12B(\lambda).
\]
\end{proposition}

\begin{proof}
Represent a cusp of small angle \(\varphi\) by a deformation of the straight
line for which
\[
  X^1(t)=\varphi t\,\Theta(t)
\]
between the scales \(\epsilon\) and \(R\), with a smooth interpolation outside
that window.  Different smoothings change only power divergences and finite
terms.  In the logarithmic momentum window \(R^{-1}\ll |\omega|\ll
\epsilon^{-1}\),
\[
  \widehat X^1(\omega)
  =
  -\frac{\varphi}{(\omega+\ii0)^2}
  +\text{endpoint and zero-frequency terms}
  +O(R^{-1},\epsilon).
\]
The endpoint terms depend on the infrared smoothing and give only power
divergences or finite oscillatory contributions, while the polynomial/contact
terms and the \(\omega=0\) terms do not affect the coefficient of
\(\log(R/\epsilon)\).  Hence the universal logarithmic contribution is
\[
  \frac12
  \int_{R^{-1}<|\omega|<\epsilon^{-1}}
  \frac{\dd\omega}{2\pi}\,
  \frac{\pi C_D}{6}|\omega|^3
  \frac{\varphi^2}{\omega^4}
  =
  \frac{C_D}{12}\varphi^2\log\frac{R}{\epsilon}.
\]
By the definition of the cusp anomalous dimension,
\[
  \log\langle W_{\rm cusp}\rangle_{\rm ren}\big|_{\log}
  =
  -\Gamma_{\rm cusp}(\varphi)\log\frac{R}{\epsilon}.
\]
Using
\(\Gamma_{\rm cusp}(\varphi)=-B(\lambda)\varphi^2+O(\varphi^4)\) and comparing
the quadratic logarithmic coefficient gives \(B=C_D/12\).
\end{proof}

Thus the radiated power of a slowly accelerating heavy probe is
\[
  \frac{\dd E_{\rm rad}}{\dd t}=2\pi B(\lambda)\,\dot v^2 .
\]
The protected circular-loop matrix model is derived in
Theorem~\ref{thm:planar-circular-wilson-bessel}.  In the convention where
the circular-loop Gaussian coupling of
Definition~\ref{def:circular-wilson-loop-matrix-coupling} is the planar
coupling \(\lambda\) used in this chapter, the supersymmetric Ward identity
relating the circular loop to the small-angle cusp gives
\[
  B(\lambda)
  =
  \frac{\sqrt\lambda}{4\pi^2}
  \frac{I_2(\sqrt\lambda)}{I_1(\sqrt\lambda)}
  \qquad (N=\infty).
\]
The algebraic step from the circular-loop answer
\(2I_1(\sqrt\lambda)/\sqrt\lambda\) to this expression is
Remark~\ref{rem:bremsstrahlung-derivative-algebra}.  The genuinely
field-theoretic input is the protected Ward identity
\eqref{eq:bremsstrahlung-circular-loop-ward}; the QSC and mirror-TBA
descriptions of the cusped line reproduce the same function with
cusp-specific asymptotic data, providing a stringent normalization check on
\(g=\sqrt\lambda/(4\pi)\).

\paragraph{Weak and strong expansions of \(B(\lambda)\).}
Let \(z=\sqrt\lambda\) and use the planar convention
\[
  B(\lambda)
  =
  \frac{z}{4\pi^2}\frac{I_2(z)}{I_1(z)} .
\]
At weak coupling,
\begin{equation}
  B(\lambda)
  =
  \frac{\lambda}{16\pi^2}
  -\frac{\lambda^2}{384\pi^2}
  +\frac{\lambda^3}{6144\pi^2}
  -\frac{\lambda^4}{92160\pi^2}
  +O(\lambda^5).
  \label{eq:bremsstrahlung-weak-lambda}
\end{equation}
Equivalently, in the spectral-problem coupling
\(g=\sqrt\lambda/(4\pi)\),
\begin{equation}
  B(g)
  =
  g^2-\frac{2\pi^2}{3}g^4+\frac{2\pi^4}{3}g^6
  -\frac{32\pi^6}{45}g^8+O(g^{10}).
  \label{eq:bremsstrahlung-weak-g}
\end{equation}
At strong coupling,
\begin{equation}
  B(\lambda)
  =
  \frac{\sqrt\lambda}{4\pi^2}
  -\frac{3}{8\pi^2}
  +\frac{3}{32\pi^2\sqrt\lambda}
  +\frac{3}{32\pi^2\lambda}
  +\frac{63}{512\pi^2\lambda^{3/2}}
  +O(\lambda^{-2}).
  \label{eq:bremsstrahlung-strong-lambda}
\end{equation}

The weak expansion follows from the defining power series
\[
  I_\nu(z)
  =
  \sum_{k=0}^{\infty}
  \frac{1}{k!\,\Gamma(k+\nu+1)}
  \left(\frac z2\right)^{2k+\nu}.
\]
For the required orders,
\[
\begin{aligned}
  I_1(z)
  &=
  \frac z2+\frac{z^3}{16}+\frac{z^5}{384}
  +\frac{z^7}{18432}+O(z^9),\\
  I_2(z)
  &=
  \frac{z^2}{8}+\frac{z^4}{96}+\frac{z^6}{3072}
  +\frac{z^8}{184320}+O(z^{10}).
\end{aligned}
\]
Dividing the two series gives
\[
  \frac{I_2(z)}{I_1(z)}
  =
  \frac z4-\frac{z^3}{96}+\frac{z^5}{1536}
  -\frac{z^7}{23040}+O(z^9).
\]
Multiplication by \(z/(4\pi^2)\) and \(z^2=\lambda\) gives
\eqref{eq:bremsstrahlung-weak-lambda}; substituting
\(\lambda=16\pi^2g^2\) gives \eqref{eq:bremsstrahlung-weak-g}.

For the strong expansion use the large-positive-\(z\) asymptotic expansion
\begin{equation}
  I_\nu(z)
  \sim
  \frac{\ee^z}{\sqrt{2\pi z}}
  \sum_{k=0}^{\infty}
  \frac{(-1)^k}{k!(8z)^k}
  \prod_{j=1}^{k}\bigl(4\nu^2-(2j-1)^2\bigr).
  \label{eq:modified-bessel-large-z-asymptotic}
\end{equation}
Let us recall the coefficient derivation, since the signs are a common
source of normalization errors.  The modified Bessel equation is
\[
  I_\nu''(z)+\frac1z I_\nu'(z)
  -\left(1+\frac{\nu^2}{z^2}\right)I_\nu(z)=0 .
\]
Insert
\[
  I_\nu(z)\sim \frac{\ee^z}{\sqrt{2\pi z}}
  S_\nu(z),\qquad
  S_\nu(z)=\sum_{k=0}^{\infty}a_k^{(\nu)}z^{-k},
  \qquad a_0^{(\nu)}=1 .
\]
After cancellation of the leading exponential part, \(S_\nu\) must satisfy
formally
\[
  2S_\nu'(z)+S_\nu''(z)
  +\left(\frac14-\nu^2\right)z^{-2}S_\nu(z)=0 .
\]
Equating the coefficient of \(z^{-k-1}\), for \(k\geq 1\), gives
\[
  a_k^{(\nu)}
  =
  -\frac{4\nu^2-(2k-1)^2}{8k}\,
  a_{k-1}^{(\nu)} ,
\]
which is precisely \eqref{eq:modified-bessel-large-z-asymptotic}.  The
classical asymptotic theorem for this ordinary differential equation gives
the corresponding remainder after truncation on every closed angular
subsector containing the positive real axis and not crossing a Stokes ray; in
this proposition only the positive real \(z=\sqrt\lambda\) ray is used.
The common factor \(\ee^z/\sqrt{2\pi z}\) cancels in the ratio.  For
\(\nu=1,2\),
\[
\begin{aligned}
  I_1(z)&\sim
  \frac{\ee^z}{\sqrt{2\pi z}}
  \left(
  1-\frac{3}{8z}-\frac{15}{128z^2}
  -\frac{315}{3072z^3}-\frac{4725}{32768z^4}
  +O(z^{-5})
  \right),\\
  I_2(z)&\sim
  \frac{\ee^z}{\sqrt{2\pi z}}
  \left(
  1-\frac{15}{8z}+\frac{105}{128z^2}
  +\frac{945}{3072z^3}+\frac{10395}{32768z^4}
  +O(z^{-5})
  \right).
\end{aligned}
\]
Series division gives
\[
  \frac{I_2(z)}{I_1(z)}
  =
  1-\frac{3}{2z}+\frac{3}{8z^2}
  +\frac{3}{8z^3}+\frac{63}{128z^4}
  +O(z^{-5}).
\]
Multiplying by \(z/(4\pi^2)\) gives
\eqref{eq:bremsstrahlung-strong-lambda}.

These expansions describe a near-BPS Wilson-line observable.  The lightlike
cusp coefficient of Section~\ref{sec:qcd-cusped-wilson-lines-large-spin} and
the planar large-spin scaling function \(f(g)\) of
Chapter~\ref{ch:planar-n4-asymptotic-bethe-ansatz} are cusp data in different
analytic limits of the generalized cusp.  Their common normalization check is
the line-defect definition \(C_D=12B(\lambda)\) together with the protected
circular-loop Ward identity.

\section{Wilson-Line Flux Tube as a Conformal Spectral Datum}
\label{sec:planar-n4-wilson-line-flux-tube-problem}

Pure Yang--Mills glueball spectroscopy, \(\mathcal N=1\) SYM chiral branches,
softly broken \(\mathcal N=2\) monopole condensation, and planar
\(\mathcal N=4\) Wilson-line integrability are related examples of
gauge-theory spectral data, but the objects being measured differ.  In
planar \(\mathcal N=4\) SYM, conformal invariance supplies no intrinsic mass
gap.  The analog of a flux-tube spectrum is instead a defect spectrum: the
Hamiltonian is the generator of translations along a line or of boosts around
a null cusp, and the excitations are states of the defect Hilbert space
associated to the Wilson line.

\begin{definition}[Planar \(\mathcal N=4\) Wilson-line flux-tube datum]
\label{def:planar-n4-wilson-line-flux-tube-problem}
Fix the planar limit of \(\mathcal N=4\) \(SU(N)\) SYM with
\(\lambda=g_{\rm YM}^2N\) and \(g=\sqrt\lambda/(4\pi)\).  A Wilson-line
flux-tube datum consists of:
\begin{enumerate}[label=(\roman*)]
\item a half-BPS straight Maldacena--Wilson line or its analytic continuation
  to a null cusp, with the scalar coupling normalized as in the preceding
  section;
\item the defect Hilbert space \(\mathcal H_{\rm def}\) obtained by radial or
  boost quantization around the line or cusp, with Hamiltonian \(H_{\rm def}\)
  and spatial momentum \(P_{\rm def}\) along the defect;
\item local defect fields \(\mathbb O_a(t)\), including the displacement
  operators \(\mathbb D_i(t)\), scalar insertions, fermion insertions, and
  covariant derivatives along the line;
\item a spectral resolution of connected defect two-point functions,
  \begin{equation}
    \langle\mathbb O_a(t)\mathbb O_b(0)\rangle_{\rm conn}
    =
    \sum_\alpha
    \int d\mu_\alpha(u)\,
    F_{a,\alpha}(u)F_{b,\alpha}(u)^*
    \ee^{-E_\alpha(u)|t|+\ii p_\alpha(u)t},
    \label{eq:planar-n4-defect-spectral-resolution}
  \end{equation}
  where \(u\) denotes rapidity-type spectral coordinates when the
  integrability description is available;
\item transition data, such as pentagon form factors or mirror-TBA kernels,
  that reconstruct multiparticle defect correlators from the one-particle
  energies, momenta, measures, and scattering phases.
\end{enumerate}
The cusp anomalous dimension is the defect vacuum energy density in the
null-cusp frame, while the generalized scaling function and pentagon OPE data
are higher spectral coordinates of the same defect system.
\end{definition}

\begin{remark}[Cusp anomalous dimension as defect ground-state energy]
\label{rem:cusp-as-defect-ground-state-energy}
Assume the null-cusp Wilson line admits a defect Hamiltonian description on a
strip of length \(L=\log(R/\epsilon)\), where \(R\) and \(\epsilon\) are the
infrared and ultraviolet scales used in the cusp renormalization.  If
\(E_0(\varphi,\theta)\) denotes the extensive ground-state energy in this
strip normalization, then
\begin{equation}
  E_0(\varphi,\theta)
  =
  \Gamma_{\rm cusp}(\varphi,\theta)\,L+O(1).
  \label{eq:cusp-ground-state-energy}
\end{equation}
For the lightlike cusp, the coefficient of \(L\) is the lightlike cusp
anomalous dimension \(\Gamma_{\rm cusp}(\lambda)\).  In the large-spin
operator problem, the same coefficient controls the leading logarithmic
growth
\begin{equation}
  \Delta(S)-S
  =
  \Gamma_{\rm cusp}(\lambda)\log S+O(S^0)
  \qquad (S\to\infty)
  \label{eq:large-spin-cusp-growth}
\end{equation}
in the normalization of
Chapter~\ref{ch:planar-n4-asymptotic-bethe-ansatz}.

The renormalized cusped-line expectation value has the defining logarithmic
form
\[
  \log\langle W_{\rm cusp}\rangle_{\rm ren}
  =
  -\Gamma_{\rm cusp}(\varphi,\theta)\log\frac{R}{\epsilon}+O(1).
\]
In the defect Hamiltonian representation, the same path integral on the
logarithmic strip is
\[
  \langle W_{\rm cusp}\rangle_{\rm ren}
  =
  \langle B_{\rm out}|\ee^{-LH_{\rm def}}|B_{\rm in}\rangle_{\rm ren},
  \qquad
  L=\log\frac{R}{\epsilon},
\]
with boundary states determined by the ultraviolet and infrared
regularizations.  If the lowest defect energy is isolated in the sector
selected by the cusp boundary conditions, the large-\(L\) asymptotic is
\[
  \log\langle W_{\rm cusp}\rangle_{\rm ren}
  =
  -E_0(\varphi,\theta)+O(L^0)
  =
  -\varepsilon_0(\varphi,\theta)L+O(L^0),
\]
where \(E_0=\varepsilon_0L\) is extensive in the strip length.  Comparing the
coefficients of \(L\) gives
\(\varepsilon_0=\Gamma_{\rm cusp}\), which is
\eqref{eq:cusp-ground-state-energy}.  The large-spin relation
\eqref{eq:large-spin-cusp-growth} follows from the same logarithmic
renormalization of the lightlike Wilson line that appears in the factorized
soft-collinear limit of twist-two operators; in the planar integrability
description this coefficient is reproduced by the BES equation and by the
large-spin limit of the asymptotic Bethe equations.
\end{remark}

\subsection{Pentagon Channel Expansion as a Defect Spectral Resolution}

The pentagon OPE used for null polygonal Wilson loops is a spectral
resolution in the defect Hilbert space of the null-cusp flux tube.  The word
``pentagon'' does not replace the Hilbert-space statement: it names the
transition amplitude between two adjacent channel decompositions of the same
renormalized Wilson-line observable.

Fix a channel of a null polygon whose conformal cross-ratios are parametrized
by \((\tau,\sigma,\phi)\).  Here \(\tau\) is the Euclidean propagation length
in the chosen flux-tube channel, \(\sigma\) is the translation coordinate
along the defect, and \(\phi\) is the angle conjugate to the transverse
rotation charge.  The channel evolution operator is
\begin{equation}
  U(\tau,\sigma,\phi)
  =
  \exp\{-\tau H_{\rm def}
        +\ii\sigma P_{\rm def}
        +\ii\phi J_{\rm def}\},
  \label{eq:n4-flux-tube-channel-evolution}
\end{equation}
where \(H_{\rm def}\), \(P_{\rm def}\), and \(J_{\rm def}\) are the energy,
longitudinal momentum, and transverse angular-momentum generators of
Definition~\ref{def:planar-n4-wilson-line-flux-tube-problem}.  Let
\(\langle\mathcal P_{\rm R}|\) and \(|\mathcal P_{\rm L}\rangle\) denote the
right and left boundary states prepared by the adjacent Wilson-line segments
in this channel, after the local cusp and edge renormalizations have been
fixed.  The channel contribution is the matrix element
\begin{equation}
  \mathcal W_{\rm ch}(\tau,\sigma,\phi)
  =
  \langle\mathcal P_{\rm R}|
  U(\tau,\sigma,\phi)
  |\mathcal P_{\rm L}\rangle .
  \label{eq:n4-pentagon-channel-matrix-element}
\end{equation}

\paragraph{Pentagon OPE from the defect spectral resolution.}
Assume the planar Wilson-line defect datum of
Definition~\ref{def:planar-n4-wilson-line-flux-tube-problem} and assume that,
in the channel under consideration, the identity on \(\mathcal H_{\rm def}\)
admits the generalized multiparticle resolution
\begin{equation}
  \mathbf 1_{\rm def}
  =
  |0\rangle\langle0|
  +
  \sum_{k\geq1}
  \sum_{a_1,\ldots,a_k}
  \frac{1}{|\operatorname{Aut}(a_1,\ldots,a_k)|}
  \int
  \prod_{r=1}^k
  \frac{\dd u_r}{2\pi}\,\mu_{a_r}(u_r)\,
  |\mathbf a,\mathbf u\rangle\langle\mathbf a,\mathbf u| ,
  \label{eq:n4-flux-tube-identity-resolution}
\end{equation}
where \(a_r\) labels the one-particle species, \(u_r\) is its rapidity
coordinate, \(\mu_a(u)\) is the one-particle spectral measure in this
normalization, and
\(|\operatorname{Aut}(a_1,\ldots,a_k)|\) is the product of factorials of the
multiplicities of identical species.  Suppose further that these states
diagonalize \(H_{\rm def}\), \(P_{\rm def}\), and \(J_{\rm def}\) additively:
\[
  E(\mathbf a,\mathbf u)=\sum_{r=1}^kE_{a_r}(u_r),\quad
  p(\mathbf a,\mathbf u)=\sum_{r=1}^kp_{a_r}(u_r),\quad
  m(\mathbf a)=\sum_{r=1}^km_{a_r}.
\]
Then \eqref{eq:n4-pentagon-channel-matrix-element} has the expansion
\begin{align}
  \mathcal W_{\rm ch}(\tau,\sigma,\phi)
  &=
  \langle\mathcal P_{\rm R}|0\rangle
  \langle0|\mathcal P_{\rm L}\rangle
  \notag\\
  &\quad+
  \sum_{k\geq1}
  \sum_{a_1,\ldots,a_k}
  \frac{1}{|\operatorname{Aut}(a_1,\ldots,a_k)|}
  \int
  \prod_{r=1}^k
  \frac{\dd u_r}{2\pi}\,\mu_{a_r}(u_r)\,
  \notag\\
  &\qquad\times
  \exp\!\left[
    -\tau\sum_{r=1}^kE_{a_r}(u_r)
    +\ii\sigma\sum_{r=1}^kp_{a_r}(u_r)
    +\ii\phi\sum_{r=1}^km_{a_r}
  \right]
  \notag\\
  &\qquad\times
  P_{\rm R}(\mathbf a,\mathbf u)\,
  P_{\rm L}(\mathbf a,\mathbf u)^* ,
  \label{eq:n4-pentagon-ope-spectral-form}
\end{align}
where
\[
  P_{\rm R}(\mathbf a,\mathbf u)
  =
  \langle\mathcal P_{\rm R}|\mathbf a,\mathbf u\rangle,
  \qquad
  P_{\rm L}(\mathbf a,\mathbf u)
  =
  \langle\mathcal P_{\rm L}|\mathbf a,\mathbf u\rangle .
\]

Insert \eqref{eq:n4-flux-tube-identity-resolution} between
\(\langle\mathcal P_{\rm R}|\) and
\(U(\tau,\sigma,\phi)|\mathcal P_{\rm L}\rangle\) in
\eqref{eq:n4-pentagon-channel-matrix-element}.  On the vacuum term,
\(H_{\rm def}\), \(P_{\rm def}\), and \(J_{\rm def}\) give the chosen vacuum
normalization.  On a multiparticle eigenstate, the assumed additivity gives
\[
  U(\tau,\sigma,\phi)|\mathbf a,\mathbf u\rangle
  =
  \exp\!\left[
    -\tau\sum_rE_{a_r}(u_r)
    +\ii\sigma\sum_rp_{a_r}(u_r)
    +\ii\phi\sum_rm_{a_r}
  \right]
  |\mathbf a,\mathbf u\rangle .
\]
The overlap with the right and left boundary states gives the two transition
amplitudes displayed in
\eqref{eq:n4-pentagon-ope-spectral-form}.  The automorphism factor is the
standard quotient by permutations of indistinguishable equal-species
particles; without it, the same unordered spectral state would be integrated
once for each such permutation.  This proves the expansion.

In the integrability solution, the overlaps
\(P_{\rm R}\) and \(P_{\rm L}\) are encoded by elementary pentagon transition
functions.  For a diagonal scalar component of two ordered one-particle
excitations in a fixed rapidity normalization, Watson consistency is the
identity
\begin{equation}
  P_{a|b}(u|v)
  =
  S_{ab}(u,v)\,P_{b|a}(v|u),
  \label{eq:n4-pentagon-watson-diagonal}
\end{equation}
with \(S_{ab}\) the flux-tube scattering phase or matrix element in the same
rapidity normalization.  Mirror moves relate a particle crossing an edge of
the pentagon to an excitation in the adjacent channel.  These relations are
powerful constraints, but they are not substitutes for the spectral datum:
the species, rapidity sheets, measures, transition normalizations, and
renormalized Wilson-line observable must all be fixed before the formula has
a mathematical meaning.

This defect datum is the \(\mathcal N=4\) member of the broader
gauge-theory family discussed in
Section~\ref{sec:pure-sym-protected-spectral-data}: the protected local
operators, line operators, and spectral measures must be defined before
their integrable or semiclassical solution methods are invoked.  The
integrability problem in planar \(\mathcal N=4\) SYM gives an unusually
complete solution of the defect spectral data; a derivation from a
nonperturbatively defined four-dimensional planar QFT remains part of the
proof boundary below.

\section{Proof Boundary}

Planar integrability gives a highly constrained exact spectral structure for
four-dimensional interacting QFT.  Symmetry fixes matrix scattering, crossing
fixes the scalar factor, Bethe equations organize asymptotic spectra, mirror
TBA incorporates finite size, and the QSC compresses the exact spectral
problem into analytic functional equations.  The remaining mathematical task
is to derive this structure from a precise nonperturbative formulation of
planar \(\mathcal N=4\) SYM, including the topology in which local operators,
large-\(N\) limits, and analytic continuation in \(\lambda\) are controlled.
