\chapter{Spectral Bridges Among Supersymmetric Yang--Mills Cousins}
\label{ch:susy-yang-mills-family-spectra}

\ConventionsLorentzian

The preceding chapters developed three separate structures: pure
\(\mathcal N=1\) Yang--Mills chiral dynamics, pure \(\mathcal N=2\)
Seiberg--Witten special geometry, and planar \(\mathcal N=4\) spectral
integrability.  They are connected by relevant deformations, but the
connection has several layers.  Holomorphic quantities can often be matched
by scale coordinates and protected Ward identities.  Massive spectra,
confining strings, and line-operator Hilbert spaces require additional
nonperturbative data.  This chapter defines the comparison problem.

\section{Comparison Problem}
\label{sec:susy-ym-family-comparison-problem}

Let \(G=SU(N_c)\) with \(N_c\ge2\).  A comparison among the
\(\mathcal N=4,\mathcal N=2,\mathcal N=1\), and bosonic Yang--Mills theories
is a comparison among four QFT problems, not a single family of Lagrangian
symbols.  The amount of supersymmetry determines which quantities are
protected, which deformations preserve a chiral coordinate chart, and which
spectral questions remain dynamical.

\begin{definition}[Yang--Mills cousin comparison problem]
\label{def:ym-cousin-comparison-problem}
A Yang--Mills cousin comparison problem is specified by:
\begin{enumerate}[label=(\roman*)]
\item a gauge group \(G=SU(N_c)\) or a specified global form with a specified
  line-operator lattice;
\item a UV description of the \(\mathcal N=4\) theory in the
  \(\mathcal N=1\) field variables
  \[
    V,\qquad \Phi_1,\Phi_2,\Phi_3\in\operatorname{ad}G,
  \]
  with holomorphic coupling
  \[
    \tau=\frac{\theta}{2\pi}+\frac{4\pi\ii}{g_{\rm YM}^2}
  \]
  in the trace convention fixed in Chapter~\ref{ch:susy-gauge-theory};
\item a finite list of relevant deformation parameters, for example adjoint
  chiral masses \(m_i\), an \(\mathcal N=2\to\mathcal N=1\) mass coordinate
  \(\mu\), and a soft gaugino-mass spurion \(m_\lambda\);
\item for each deformation regime, a declared set of observables:
  gauge-invariant local operators, center-charged line operators, domain-wall
  boundary conditions, and defect operators on Wilson lines;
\item a status label for each comparison: holomorphic scale matching,
  protected \(F\)-term matching, semiclassical Abelian approximation, lattice
  spectral extraction, planar-integrability reconstruction, or open
  nonperturbative construction.
\end{enumerate}
Two rows of the comparison are compared only after the observable type and
status label have been fixed.
\end{definition}

The basic ladder is:
\[
\begin{array}{ccl}
\mathcal N=4
&\xrightarrow{\;m_{\rm hyper}\;}
&\mathcal N=2^\ast
\xrightarrow{\;E\ll |m_{\rm hyper}|\;}
\text{pure }\mathcal N=2,\\[3pt]
\mathcal N=4
&\xrightarrow{\;m_1,m_2,m_3\;}
&\mathcal N=1^\ast
\xrightarrow{\;E\ll \min_i|m_i|\;}
\text{pure }\mathcal N=1,\\[3pt]
\text{pure }\mathcal N=2
&\xrightarrow{\;\mu\,u\;}
&\text{softly broken }\mathcal N=1\text{ with Abelian confinement},\\[3pt]
\text{pure }\mathcal N=1
&\xrightarrow{\;X_\lambda S+\mathrm{c.c.}\;}
&\text{bosonic Yang--Mills deformation problem}.
\end{array}
\]
The arrows are deformation regimes.  A continuum construction for the
endpoint theory, or a proof that a path of deformations reaches a chosen
endpoint without crossing a phase boundary in the observables of interest, is
extra information.

\section{Holomorphic Scale Matching}
\label{sec:susy-ym-family-scale-matching}

Holomorphic scale matching is the cleanest part of the comparison.  It is a
statement about chiral source coordinates and threshold matching in a
holomorphic Wilsonian scheme.  It does not compute the nonchiral spectrum.

\begin{definition}[Holomorphic threshold matching]
\label{def:holomorphic-threshold-matching}
Let a four-dimensional supersymmetric gauge theory with gauge algebra
\(\mathfrak{su}(N_c)\) have holomorphic one-loop coefficient \(b_{\rm high}\)
above a chiral mass threshold \(m\) and coefficient \(b_{\rm low}\) below the
threshold.  A holomorphic threshold matching at the scale \(m\) is the
identity of chiral scale coordinates
\begin{equation}
  \Lambda_{\rm low}^{b_{\rm low}}
  =
  m^{\,b_{\rm low}-b_{\rm high}}\,
  \Lambda_{\rm high}^{b_{\rm high}},
  \label{eq:holomorphic-threshold-matching}
\end{equation}
up to finite holomorphic scheme rescalings already included in the
definitions of \(m\) and \(\Lambda\).
\end{definition}

Write \(\mathcal N=4\) Yang--Mills in \(\mathcal N=2\) variables as one
\(\mathcal N=2\) vector multiplet plus one adjoint hypermultiplet.  Give the
hypermultiplet a holomorphic mass \(m_{\rm h}\).  In the region
\[
  E\ll |m_{\rm h}|,
\]
the remaining pure \(\mathcal N=2\) holomorphic scale satisfies
\begin{equation}
  \Lambda_{\mathcal N=2}^{2N_c}
  =
  m_{\rm h}^{2N_c}\,q,
  \qquad
  q=\ee^{2\pi\ii\tau}.
  \label{eq:n4-to-n2-scale-matching}
\end{equation}
The \(\mathcal N=4\) theory has vanishing holomorphic beta coefficient, so
the high-energy scale coordinate is the dimensionless instanton factor
\(q=\ee^{2\pi\ii\tau}\).  Pure \(\mathcal N=2\) \(SU(N_c)\) Yang--Mills has
holomorphic coefficient
\[
  b_{\mathcal N=2}=2N_c .
\]
Applying \eqref{eq:holomorphic-threshold-matching} with
\[
  b_{\rm high}=0,\qquad b_{\rm low}=2N_c,\qquad m=m_{\rm h}
\]
gives \eqref{eq:n4-to-n2-scale-matching}.  The dimension of the right-hand
side is \(2N_c\), matching the dimension of
\(\Lambda_{\mathcal N=2}^{2N_c}\).

Write \(\mathcal N=4\) Yang--Mills in \(\mathcal N=1\) variables as one
vector multiplet and three adjoint chiral multiplets
\(\Phi_1,\Phi_2,\Phi_3\).  Give the three adjoint chiral multiplets
holomorphic masses \(m_1,m_2,m_3\).  In the region
\[
  E\ll \min_i |m_i|,
\]
the pure \(\mathcal N=1\) holomorphic scale satisfies
\begin{equation}
  \Lambda_{\mathcal N=1}^{3N_c}
  =
  (m_1m_2m_3)^{N_c}\,q,
  \qquad
  q=\ee^{2\pi\ii\tau}.
  \label{eq:n4-to-n1-scale-matching}
\end{equation}
The high-energy \(\mathcal N=4\) coefficient is zero.  Pure
\(\mathcal N=1\) \(SU(N_c)\) Yang--Mills has coefficient
\[
  b_{\mathcal N=1}=3N_c .
\]
Each massive adjoint chiral multiplet changes the holomorphic coefficient by
\(T(\operatorname{ad})=N_c\).  Applying
\eqref{eq:holomorphic-threshold-matching} successively at the three mass
thresholds gives
\[
  \Lambda_{\mathcal N=1}^{3N_c}
  =
  m_1^{N_c}m_2^{N_c}m_3^{N_c}q,
\]
which is \eqref{eq:n4-to-n1-scale-matching}.  The mass dimension is
\(3N_c\), as required for the pure \(\mathcal N=1\) scale coordinate.

\begin{example}[\(\mathcal N=1^\ast\) chiral critical equations]
\label{ex:n1-star-critical-equations}
In an \(\mathcal N=1\) presentation, the equal-mass \(\mathcal N=1^\ast\)
superpotential is
\begin{equation}
  W
  =
  \sqrt2\,\operatorname{tr}\Phi_1[\Phi_2,\Phi_3]
  +
  \frac{m}{2}\sum_{i=1}^3\operatorname{tr}\Phi_i^2 .
  \label{eq:n1-star-superpotential}
\end{equation}
The \(F\)-term equations are
\begin{equation}
  \sqrt2[\Phi_2,\Phi_3]+m\Phi_1=0,
  \qquad
  \sqrt2[\Phi_3,\Phi_1]+m\Phi_2=0,
  \qquad
  \sqrt2[\Phi_1,\Phi_2]+m\Phi_3=0.
  \label{eq:n1-star-f-terms}
\end{equation}
If \(J_i\) are matrices obeying
\begin{equation}
  [J_i,J_j]=\ii\epsilon_{ijk}J_k,
  \label{eq:su2-fuzzy-sphere-algebra}
\end{equation}
then
\begin{equation}
  \Phi_i=\frac{\ii m}{\sqrt2}\,J_i
  \label{eq:n1-star-fuzzy-sphere-solution}
\end{equation}
solves \eqref{eq:n1-star-f-terms}.  Such solutions organize chiral vacua of
the massive adjoint theory by embeddings of \(\mathfrak{su}(2)\) into
\(\mathfrak{su}(N_c)\).  The full massive spectrum and confinement data also
depend on the nonchiral dynamics and on the chosen line-operator lattice.
\end{example}

\section{What Is Uniform Across the Ladder}
\label{sec:susy-ym-family-uniform-data}

The uniform part of the ladder is a map of observables, not a universal
numerical formula.  Four classes appear repeatedly.

\begin{definition}[Family observable comparison table]
\label{def:susy-ym-family-observable-table}
For a Yang--Mills cousin comparison problem, the repeated observable classes
are the following.
\begin{center}
\small
\begin{tabular}{p{0.22\textwidth}|p{0.39\textwidth}|p{0.28\textwidth}}
\textbf{observable type} & \textbf{representative object} & \textbf{status}\\
\hline
chiral local &
  \(S,\ u,\ \operatorname{tr}\Phi^k\) &
  holomorphic or protected\\
massive local spectral &
  \(\rho_{ij}(B)\) for gauge-invariant sources &
  spectral measure; lattice extraction\\
center-charged line &
  \(W_R(C),\ H_m(C)\), flux-sector transfer matrix &
  declared line-sector problem\\
defect spectral &
  \(\mathcal H_{\rm def},\ H_{\rm def},\mathbb O_a(t)\) &
  Wilson-line or cusp spectral problem
\end{tabular}
\end{center}
where \(S\) is the pure \(\mathcal N=1\) glueball coordinate of
Section~\ref{sec:pure-sym-gaugino-condensation}, \(u\) is the rank-one
\(\mathcal N=2\) Coulomb coordinate of
Chapter~\ref{ch:susy-four-dimensional-n2-seiberg-witten}, \(\rho_{ij}\) is
the spectral measure of
Definition~\ref{def:pure-sym-vacuum-spectral-problem}, \(W_R(C)\) is a Wilson
line in representation \(R\), \(H_m(C)\) is a magnetic line with magnetic
charge \(m\), and \(\mathcal H_{\rm def}\) is the defect Hilbert space of
Definition~\ref{def:planar-n4-wilson-line-flux-tube-problem}.
\end{definition}

\begin{controlledapproximation}[Observable-type comparison across a controlled threshold]
\label{ca:observable-type-threshold-comparison}
Consider a relevant mass deformation with threshold \(M\) between two
supersymmetric Yang--Mills cousins.  The following comparison is meaningful
only after the threshold problem is supplied with:
\begin{enumerate}[label=(\roman*)]
\item a Wilsonian description exists above and below \(M\) in a common
  regulator or in compatible holomorphic schemes;
\item the local source under consideration has external momenta
  \(|p_a|\ll |M|\);
\item every massive field integrated out has a gap bounded below by
  \(c|M|\) for some \(c>0\) in the chosen background;
\item the line-operator sector is kept fixed by declaring the global form and
  allowed electric and magnetic charges before the deformation.
\end{enumerate}
Under these hypotheses, chiral local sources match by holomorphic threshold
coordinates, local nonchiral correlators match after adding local
higher-dimension Wilsonian operators, and center-charge labels of line
operators are transported as declared labels.  The spectrum is a separate
dynamical problem: numerical masses, string tensions, and line-defect
excitation energies are functions of the deformation parameters and require
the corresponding Hilbert-space, transfer-matrix, or line-defect spectral
analysis.

For chiral local sources, the supersymmetric Wilsonian action separates the
holomorphic \(F\)-terms from \(D\)-terms.  The threshold contribution of a
massive chiral multiplet to the holomorphic gauge coupling is one-loop exact
in the Wilsonian holomorphic coordinate, giving the scale matching of
Definition~\ref{def:holomorphic-threshold-matching}.  Thus the chiral
coordinates match by finite holomorphic maps such as
\eqref{eq:n4-to-n2-scale-matching} and \eqref{eq:n4-to-n1-scale-matching}.

For nonchiral local correlators at momenta \(|p_a|\ll |M|\), the massive
propagators admit an expansion in powers of \(p/M\) after the regulator and
subtraction scheme have been fixed.  The coefficients are local functionals
of the light fields and sources, hence enter the Wilsonian action as
higher-dimension local operators.  This is an equality of low-momentum
Wilsonian coordinates at fixed order in the derivative expansion, with a
remainder controlled only when the chosen regulator supplies the corresponding
analytic estimates.

Line operators require the global form and charge lattice as part of the
definition.  A mass threshold changes the dynamics of screening and
confinement, but it does not change the declared lattice of allowed line
charges.  Therefore the center or charge label can be transported through the
comparison problem.  The area-law coefficient or defect excitation energy is
obtained from the Hilbert-space or transfer-matrix problem in the deformed
theory, so it remains a dynamical spectral quantity.
\end{controlledapproximation}

\section[Center Strings and k-String Tension Comparisons]{Center Strings and \(k\)-String Tension Comparisons}
\label{sec:susy-ym-family-k-string-tensions}

For \(SU(N_c)\) with the usual Wilson lines, the center charge of a
representation is an element of \(\mathbb Z_{N_c}\).  A confining theory may
have stable flux sectors labeled by
\[
  k\in\{0,1,\ldots,N_c-1\},
\]
with \(k\) and \(N_c-k\) related by charge conjugation.

\begin{definition}[\(k\)-string tension extraction]
\label{def:k-string-tension-extraction}
A \(k\)-string tension extraction for a confining Yang--Mills theory is
specified by:
\begin{enumerate}[label=(\roman*)]
\item a regulator or continuum Hilbert-space construction with spatial
  transfer matrix \(T_L\) on a box whose long direction has length \(L\);
\item a center-flux sector \(k\in\mathbb Z_{N_c}\);
\item the lowest energy \(E_k(L)\) in that sector after the vacuum energy is
  subtracted;
\item a limit
  \begin{equation}
    T_k=\lim_{L\to\infty}\frac{E_k(L)}{L}
    \label{eq:k-string-transfer-tension}
  \end{equation}
  whenever the limit exists.
\end{enumerate}
Equivalently, in an area-law regime for a large rectangular Wilson loop in a
representation of \(N\)-ality \(k\),
\begin{equation}
  \langle W_k(C)\rangle
  =
  \exp\!\bigl(-T_k\,\operatorname{Area}(C)+o(\operatorname{Area}(C))\bigr)
  \label{eq:k-string-area-law}
\end{equation}
defines the same tension if the transfer-matrix and Wilson-loop formulations
have been proved equivalent for the regulator under consideration.
\end{definition}

Two comparison formulae are useful reference functions:
\begin{equation}
  R_k^{\rm sine}(N_c)
  =
  \frac{\sin(\pi k/N_c)}{\sin(\pi/N_c)},
  \qquad
  R_k^{\rm Casimir}(N_c)
  =
  \frac{k(N_c-k)}{N_c-1}.
  \label{eq:k-string-tension-ratios}
\end{equation}
The first appears in Abelianized supersymmetric regimes and in several
string-inspired large-\(N_c\) calculations.  The second is a quadratic
color-factor comparison coordinate.  In this chapter they are comparison
functions until a specified QFT framework supplies a derivation of a spectral
tension ratio.

For \(1\le k\le N_c-1\),
\begin{equation}
  R_k^{\rm sine}(N_c)=R_{N_c-k}^{\rm sine}(N_c),
  \qquad
  R_k^{\rm Casimir}(N_c)=R_{N_c-k}^{\rm Casimir}(N_c).
  \label{eq:k-string-charge-conjugation-symmetry}
\end{equation}
At fixed \(k\) and large \(N_c\),
\begin{equation}
  R_k^{\rm sine}(N_c)
  =
  k-\frac{\pi^2}{6}\frac{k(k^2-1)}{N_c^2}
  +O(N_c^{-4}),
  \label{eq:sine-law-large-n-expansion}
\end{equation}
while
\begin{equation}
  R_k^{\rm Casimir}(N_c)
  =
  k-\frac{k(k-1)}{N_c}+O(N_c^{-2}).
  \label{eq:casimir-large-n-expansion}
\end{equation}
These are finite comparison identities.  The symmetry follows from
\[
  \sin\!\left(\frac{\pi(N_c-k)}{N_c}\right)
  =
  \sin\!\left(\pi-\frac{\pi k}{N_c}\right)
  =
  \sin\!\left(\frac{\pi k}{N_c}\right)
\]
and from
\[
  (N_c-k)k=k(N_c-k).
\]
The sine-law expansion is obtained by dividing
\[
  \sin\frac{\pi k}{N_c}
  =
  \frac{\pi k}{N_c}
  -
  \frac{\pi^3k^3}{6N_c^3}
  +O(N_c^{-5}),
  \qquad
  \sin\frac{\pi}{N_c}
  =
  \frac{\pi}{N_c}
  -
  \frac{\pi^3}{6N_c^3}
  +O(N_c^{-5}).
\]
This gives
\[
  R_k^{\rm sine}
  =
  k\left(1-\frac{\pi^2k^2}{6N_c^2}\right)
  \left(1+\frac{\pi^2}{6N_c^2}\right)
  +O(N_c^{-4}),
\]
which is \eqref{eq:sine-law-large-n-expansion}.  For the Casimir comparison,
\[
  \frac{k(N_c-k)}{N_c-1}
  =
  k\frac{1-k/N_c}{1-1/N_c}
  =
  k\left(1-\frac{k}{N_c}\right)
  \left(1+\frac{1}{N_c}+O(N_c^{-2})\right),
\]
which gives \eqref{eq:casimir-large-n-expansion}.  The physics lies in
whether a specified gauge theory and regulator produce one of these finite
functions as a spectral ratio; the algebra above supplies only comparison
coordinates.

\section{Entries of the Family}
\label{sec:susy-ym-family-entries}

\subsection{Pure Bosonic Yang--Mills}

Pure bosonic Yang--Mills is the endpoint with the fewest protected
observables and the strongest need for a nonperturbative construction.  Its
glueball and flux-tube spectra are formulated in
Section~\ref{sec:lattice-ym-glueballs-flux-observables}, and their continuum
existence, mass gap, and confinement status are organized by
Conjecture~\ref{conj:pure-yang-mills-central}.  In the family comparison, the
bosonic row contains:
\[
\begin{array}{c|c}
\text{object} & \text{role}\\
\hline
\text{OS-positive lattice gauge measure} &
  \text{candidate regulator and transfer matrix}\\
\text{local operators }\operatorname{tr}F^2,\operatorname{tr}F\widetilde F,\ldots &
  J^{PC}\text{ glueball sources}\\
\text{center-flux sectors} &
  k\text{-string tensions and excited flux tubes}\\
\text{large-}N_c\text{ expansion} &
  M_{\rm glueball}=O(1),\quad \Gamma_{\rm glueball}=O(N_c^{-2})
\end{array}
\]
The bosonic theory supplies the target spectral questions listed in
Conjecture~\ref{conj:pure-yang-mills-central}: a mass gap, isolated glueball
poles or controlled spectral thresholds, stable flux sectors, and an effective
string description in the long-string regime.

\subsection[Pure N=1 Yang--Mills]{Pure \(\mathcal N=1\) Yang--Mills}

Pure \(\mathcal N=1\) Yang--Mills adds protected chiral branch information.
The continuum phase, gap, confinement, and condensate claims used here are the
central package of Conjecture~\ref{conj:pure-n1-sym-central}.  The local
coordinate
\[
  S=-\frac{1}{32\pi^2}\operatorname{tr}W^\alpha W_\alpha
\]
has branch values
\[
  \langle S\rangle_k=C\Lambda_h^3\ee^{2\pi\ii k/N_c}.
\]
The family comparison separates three outputs:
\[
\begin{array}{c|c}
\text{output} & \text{source}\\
\hline
N_c\text{ branch phases} & \mathbb Z_{2N_c}\text{ anomaly and chiral }F\text{-term}\\
T_{ij}=2|W_i-W_j| & \mathcal N=1\text{ BPS wall algebra}\\
\text{glueball/gluinoball masses} & \text{spectral measure }\rho_{ij}^{(k)}
\end{array}
\]
The first two are protected after the stated chiral-sector hypotheses and after
the central conjecture supplies the pure phases.  The third is a massive
spectral problem involving the scalar, pseudoscalar, and fermionic components
of gauge-invariant superfields.  Supersymmetry organizes isolated massive
states into multiplets when the vacuum preserves the supercharges, but it does
not supply their numerical masses.

\begin{definition}[Pure-SYM channel source table]
\label{def:pure-sym-channel-source-family}
Fix a massive branch \(k\) of pure \(\mathcal N=1\) \(SU(N_c)\) Yang--Mills
and the vacuum spectral measure of
Definition~\ref{def:pure-sym-vacuum-spectral-problem}.  Let
\[
  S=-\frac{1}{32\pi^2}\operatorname{tr}W^\alpha W_\alpha
\]
be the chiral coordinate used in
Section~\ref{sec:pure-sym-protected-spectral-data}.  The elementary source
table attached to this superfield is:
\begin{center}
\small
\begin{tabular}{p{0.20\textwidth}|p{0.37\textwidth}|p{0.32\textwidth}}
\textbf{channel} & \textbf{source family} & \textbf{spectral role}\\
\hline
bosonic even scalar &
  \(\operatorname{Re} S|_{\theta=0}\), \(\operatorname{Re}D^2S|\), and
  same-quantum-number descendants &
  overlaps with \(0^{++}\)-type atoms after operator mixing is fixed\\
bosonic odd scalar &
  \(\operatorname{Im} S|_{\theta=0}\), \(\operatorname{Im}D^2S|\), and
  same-quantum-number descendants &
  overlaps with \(0^{-+}\)-type atoms after operator mixing is fixed\\
fermionic &
  \(D_\alpha S|\), \(\overline D_{\dot\alpha}\overline S|\), and
  derivative descendants &
  overlaps with spin-\(\frac12\) atoms in the same massive multiplets\\
\end{tabular}
\end{center}
Here the vertical bar denotes restriction to \(\theta=\overline\theta=0\) after
the chosen superspace operator has been renormalized.  The table lists source
families, not particles: the spectral measure may contain several
isolated atoms and continuum components in any row, and the same atom can have
nonzero residue for several sources with the same exact quantum numbers.
Center-charged Wilson or magnetic lines belong to a separate line-sector
problem; they define flux-string sectors rather than chiral-particle masses.
\end{definition}

\paragraph{Mass degeneracy inside an isolated pure-SYM multiplet.}
Assume the vacuum branch \(k\) carries unbroken \(\mathcal N=1\) supercharges
\(Q_\alpha,\overline Q_{\dot\alpha}\) on the Hilbert space of
Definition~\ref{def:pure-sym-vacuum-spectral-problem}.  Let \(E_k(\Sigma_m)\)
be the spectral projection onto an isolated positive-mass shell
\(\Sigma_m=\{p:p^2=-m^2,\ p^0>0\}\), separated from the rest of the joint
translation spectrum in a neighborhood invariant under the little group.
Then the range of \(E_k(\Sigma_m)\) is invariant under the supercharges.  In
particular, whenever a bosonic or fermionic source in
Definition~\ref{def:pure-sym-channel-source-family} has nonzero overlap with
this atom, the complete supermultiplet generated by the supercharges has the
same mass \(m\).
The supersymmetry algebra gives
\[
  [P_\mu,Q_\alpha]=[P_\mu,\overline Q_{\dot\alpha}]=0.
\]
Therefore the mass operator
\[
  M_k^2=-\eta^{\mu\nu}P_\mu^{(k)}P_\nu^{(k)}
\]
commutes with the supercharges on their common invariant domain.  Functional
calculus then gives
\[
  Q_\alpha E_k(B)=E_k(B)Q_\alpha,\qquad
  \overline Q_{\dot\alpha}E_k(B)=E_k(B)\overline Q_{\dot\alpha}
\]
for every Borel set \(B\) of the joint spectrum for which the products are
defined by closure; applying this to the isolated shell gives invariance of
\(\operatorname{Ran}E_k(\Sigma_m)\).  Acting with the supercharges changes
spin and fermion number but not \(P^2\), so all nonzero states in the generated
irreducible massive supermultiplet have the same mass.  The statement says
nothing about the numerical value of \(m\) or about whether the atom exists;
those are spectral questions in the chosen nonperturbative construction.

\begin{definition}[Regulated pure-SYM spectral extraction]
\label{def:regulated-pure-sym-spectral-extraction}
A regulated extraction of the pure \(\mathcal N=1\) SYM spectrum consists of:
\begin{enumerate}[label=(\roman*)]
\item a gauge-field regulator and an adjoint Majorana-fermion regulator with
  a specified Pfaffian or determinant prescription;
\item a parameter \(m_{\lambda,\mathrm{reg}}\) whose tuned value defines the
  massless-gluino trajectory, together with the Ward identity or equivalent
  renormalized condition used to tune it;
\item a finite-volume transfer matrix or Euclidean reflection-positive
  substitute sufficient to define spectral estimates;
\item source matrices in the exact finite-regulator symmetry channels, with
  the continuum \(J^{PC}\) labels assigned only after the rotation and
  discrete-symmetry restoration limits have been taken;
\item a limit prescription \(a\to0\), \(L\to\infty\), and
  \(m_{\lambda,\mathrm{reg}}\to m_{\lambda,\ast}\), including the order of
  limits and the systematic error controls used in the extrapolation.
\end{enumerate}
This regulated extraction is the supersymmetric analogue of the pure-glue
construction in Section~\ref{sec:lattice-ym-glueballs-flux-observables}.  It
is not replaced by the Veneziano--Yankielowicz superpotential: the latter
supplies protected branch information, while the spectral extraction is what
can define numerical glueball,
gluinoball, and mixed supermultiplet masses.
\end{definition}

\begin{definition}[Channel-by-channel pure-SYM correlator matrices]
\label{def:pure-sym-channel-correlator-matrices}
Fix the regulated extraction of
Definition~\ref{def:regulated-pure-sym-spectral-extraction}.  Choose
zero-spatial-momentum finite-volume sources
\[
  B_{+,r},\qquad B_{-,r},\qquad F_{\alpha,r},
  \qquad r=1,\ldots,R ,
\]
where \(B_+\) has the quantum numbers of the even scalar row of
Definition~\ref{def:pure-sym-channel-source-family}, \(B_-\) has the
odd-scalar row, and \(F_\alpha\) has the fermionic row.  With the vacuum
projection removed, define the Euclidean transfer-matrix correlator matrices
\begin{align}
  C_{+}^{rs}(t)
  &=
  \langle\Omega|B_{+,r}\,\ee^{-tH_{\rm reg}}\,B_{+,s}^{\dagger}|\Omega\rangle_c,
  \label{eq:pure-sym-even-correlator-matrix}\\
  C_{-}^{rs}(t)
  &=
  \langle\Omega|B_{-,r}\,\ee^{-tH_{\rm reg}}\,B_{-,s}^{\dagger}|\Omega\rangle_c,
  \label{eq:pure-sym-odd-correlator-matrix}\\
  C_{F}^{rs}(t)
  &=
  \frac12\operatorname{tr}_{\rm rest}
  \langle\Omega|F_r\,\ee^{-tH_{\rm reg}}\,\overline F_s|\Omega\rangle_c .
  \label{eq:pure-sym-fermion-correlator-matrix}
\end{align}
The trace in \eqref{eq:pure-sym-fermion-correlator-matrix} is the declared
spin average in the rest frame; at finite lattice spacing it must be replaced
by the exact spin-little-group projection of the regulator.  The channel
matrices are not themselves a numerical result.  They specify the
finite-regulator spectral data whose leading isolated pole positions may be
compared across the even scalar, odd scalar, and fermionic channels.  The
method used to estimate those pole positions is external to the present
chapter; the pure-SYM content is the choice of channels, sources,
spin/rest-frame projections, gluino-mass tuning condition, and joint
continuum, thermodynamic, and massless-gluino limiting procedure.  A
pure-SYM spectral extraction has restored supersymmetry in a one-particle
window only after that limiting procedure makes the pole locations extracted
from the three channels agree within a stated continuum and infinite-volume
error estimate.
\end{definition}

Operationally, the pure-SYM lattice or Hamiltonian analysis should publish,
for each candidate multiplet, the three leading masses
\[
  m_+,\qquad m_-,\qquad m_F
\]
with the regulator parameters, source bases, fit windows, finite-volume
trajectory, and tuning condition for the gluino mass.  The supersymmetry
restoration diagnostic is not the equality of two names in a table; it is the
vanishing of
\begin{equation}
  \delta_{\rm SUSY}
  =
  \frac{\max\{|m_+-m_-|,|m_+-m_F|,|m_--m_F|\}}
       {\frac13(m_++m_-+m_F)}
  \label{eq:pure-sym-spectral-susy-diagnostic}
\end{equation}
in the joint continuum, infinite-volume, and massless-gluino limit, with the
limiting procedure stated as part of
Definition~\ref{def:regulated-pure-sym-spectral-extraction}.

\subsection[Pure N=2 and Softly Broken N=2]{Pure \(\mathcal N=2\) and Softly Broken \(\mathcal N=2\)}

Pure \(\mathcal N=2\) Yang--Mills has Coulomb-branch coordinate \(u\), period
vector \((a_D,a)\), and BPS central charge
\[
  Z_\gamma=n_ea+n_ma_D .
\]
At a monopole singularity, a soft \(\mathcal N=1\) deformation
\[
  W_{\mathrm{loc},\mu}
  =
  \sqrt2\,A_DM\widetilde M+\mu\,U(A_D)
\]
gives
\[
  A_D=0,\qquad
  M\widetilde M=-\frac{\mu}{\sqrt2}U'(0).
\]
The dual \(U(1)_D\) is Higgsed, and the local Abelian model supports flux
tubes carrying the original electric flux.  The Seiberg--Witten row therefore
contains a controlled Abelian confinement mechanism near the singularity.
The detailed non-Abelian \(k\)-string spectrum away from the local Abelian
regime is a separate spectral problem.

\begin{controlledapproximation}[Radial BPS string profile near a monopole point]
\label{ca:sw-radial-bps-string-profile}
Use the local Abelian-Higgs normalization of
\eqref{eq:sw-local-abelian-higgs-energy}, take \(n>0\), and choose the
orientation for which
\begin{equation}
  (D_1+\ii D_2)q=0,\qquad B_D=e_D^2(|\xi|-|q|^2).
  \label{eq:family-sw-bps-vortex-equations}
\end{equation}
With polar coordinate \((r,\varphi)\) on the plane transverse to the string,
set
\begin{equation}
  q(r,\varphi)=|\xi|^{1/2}f(R)\ee^{\ii n\varphi},
  \qquad
  A_D=n\,a(R)\,\dd\varphi,\qquad
  R=e_D|\xi|^{1/2}r .
  \label{eq:family-sw-vortex-ansatz}
\end{equation}
Finite energy and regularity impose
\begin{equation}
  f(0)=0,\qquad a(0)=0,\qquad
  f(\infty)=1,\qquad a(\infty)=1.
  \label{eq:family-sw-vortex-boundary-conditions}
\end{equation}
The first-order equations reduce to
\begin{equation}
  \frac{\dd f}{\dd R}=\frac{n}{R}(1-a)f,
  \qquad
  \frac{\dd a}{\dd R}=\frac{R}{n}(1-f^2).
  \label{eq:family-sw-vortex-radial-ode}
\end{equation}
The flux and BPS tension are
\begin{equation}
  \int_{\mathbb R^2}F_D=2\pi n,\qquad
  T_n=2\pi n|\xi|.
  \label{eq:family-sw-vortex-flux-tension}
\end{equation}
\end{controlledapproximation}

\paragraph{Derivation of the radial vortex equations.}
For the ansatz \eqref{eq:family-sw-vortex-ansatz},
\[
  D_rq=|\xi|^{1/2}e_D|\xi|^{1/2}f'(R)\ee^{\ii n\varphi},
  \qquad
  D_\varphi q=\ii n(1-a)|\xi|^{1/2}f(R)\ee^{\ii n\varphi}.
\]
Since \(D_1+\ii D_2=\ee^{\ii\varphi}(D_r+\ii r^{-1}D_\varphi)\), the first BPS
equation gives
\[
  e_D|\xi|^{1/2}f'(R)-\frac{n}{r}(1-a)f(R)=0,
\]
which is the first equation in \eqref{eq:family-sw-vortex-radial-ode}.  The
curvature of the one-form \(A_D=n\,a(R)\dd\varphi\) is
\[
  F_D=n\,a'(R)\dd R\wedge\dd\varphi,
  \qquad
  B_D=\frac{n}{r}e_D|\xi|^{1/2}a'(R).
\]
Substituting \(r=R/(e_D|\xi|^{1/2})\) into
\(B_D=e_D^2|\xi|(1-f^2)\) gives the second radial equation.  Finally,
\[
  \int_{\mathbb R^2}F_D
  =
  \int_0^{2\pi}\!\dd\varphi\int_0^\infty n\,a'(R)\dd R
  =
  2\pi n\,[a(\infty)-a(0)]
  =
  2\pi n .
\]
The tension formula is then the saturated bound
\eqref{eq:sw-local-vortex-bound} with positive orientation.  The derivation is
local in the Seiberg--Witten patch; it does not assert that a full non-Abelian
\(k\)-string spectrum is determined by the Abelian profile.

\subsection[Abelianized k-String Comparison]{Abelianized \(k\)-String Comparison}

The local \(U(1)_D\) vortex just derived is the rank-one building block.  For
\(SU(N_c)\) with \(N_c>2\), the softly broken \(\mathcal N=2\) theory has
singular loci where several mutually local magnetic states become light and
the low-energy magnetic gauge group is an Abelian torus.  The string
calculation in such a region is an Abelianized calculation.  Its output is a
precise low-energy spectral problem; comparison with the non-Abelian
pure-Yang--Mills \(k\)-string spectrum requires an additional deformation
argument.

Let \(K_{ij}\), \(1\le i,j\le N_c-1\), be the \(A_{N_c-1}\) Cartan matrix
\begin{equation}
  K_{ij}=2\delta_{ij}-\delta_{i,j+1}-\delta_{i+1,j}.
  \label{eq:an-minus-one-cartan-matrix}
\end{equation}
The simple magnetic fields of an Abelianized Seiberg--Witten patch are
labelled by these simple roots.  A soft \(\mathcal N=1\) deformation supplies
complex Fayet--Iliopoulos coordinates \(\xi_i\) for the corresponding local
Abelian Higgs factors after a magnetic duality frame has been chosen.  When
the phases of the \(\xi_i\) are aligned, the BPS square completion of the
rank-one calculation applies independently in each diagonalized magnetic
charge sector.

\begin{definition}[Abelianized \(A_{N_c-1}\) string construction]
\label{def:abelianized-a-string-approximation}
An Abelianized \(A_{N_c-1}\) string construction is specified by:
\begin{enumerate}[label=(\roman*)]
\item a magnetic low-energy patch with gauge group \(U(1)^{N_c-1}\), gauge
  fields \(A_{D,i}\), and mutually local charged monopole hypermultiplets;
\item complex FI coordinates \(\xi_i\), \(i=1,\ldots,N_c-1\), with a declared
  common phase, so that a BPS bound may be written in a single real central
  charge direction;
\item flux integers \(n_i\in\mathbb Z\) with
  \[
    \frac{1}{2\pi}\int_{\mathbb R^2}F_{D,i}=n_i;
  \]
\item a map from the Abelian magnetic flux vector \(n=(n_i)\) to the
  electric center charge \(k\in\mathbb Z_{N_c}\) of the heavy Wilson source
  being confined;
\item the domain in Coulomb-branch and deformation-parameter space in which
  the Abelian two-derivative description is used.
\end{enumerate}
This construction defines an Abelianized string spectrum in the local effective
theory.  It becomes a statement about the microscopic non-Abelian theory only
after the Seiberg--Witten effective action and the deformation to the chosen
microscopic problem have been justified in the stated regime.
\end{definition}

\paragraph{Cartan BPS bound in an aligned Abelian chart.}
Assume the Abelianized construction of
Definition~\ref{def:abelianized-a-string-approximation} and suppose that, in the
chosen magnetic frame, the static transverse energy density is a sum of
rank-one Abelian-Higgs BPS densities with aligned \(\xi_i\).  For flux vector
\(n=(n_1,\ldots,n_{N_c-1})\), every finite-energy string obeys
\begin{equation}
  T(n)\ge
  2\pi\left|\sum_{i=1}^{N_c-1}n_i\xi_i\right|.
  \label{eq:cartan-abelian-string-bps-bound}
\end{equation}
If the first-order vortex equations are solved in every occupied sector with
the common central-charge orientation, and if all nonzero \(n_i\xi_i\) have
the phase of \(\sum_j n_j\xi_j\), the bound is saturated.

For a single diagonal magnetic factor the square completion in
\eqref{eq:sw-local-abelian-higgs-energy} gives
\[
  T_i\ge 2\pi |n_i|\,|\xi_i|
\]
with equality when the flux orientation is chosen to have the same phase as
\(n_i\xi_i\).
Writing all \(\xi_i\) in the common phase direction as
\(\xi_i=\ee^{\ii\alpha}|\xi_i|\), the central charge of a multi-flux
configuration is
\[
  Z(n)=2\pi\sum_i n_i\xi_i .
\]
The triangle inequality gives
\[
  \sum_i2\pi |n_i|\,|\xi_i|
  \ge
  2\pi\left|\sum_i n_i\xi_i\right|
  =
  |Z(n)|,
\]
which is \eqref{eq:cartan-abelian-string-bps-bound}.  Saturation of the
displayed central-charge bound requires equality in the triangle inequality,
namely all nonzero \(n_i\xi_i\) have the same phase, and also saturation of
the rank-one square in every occupied sector.  The argument is a statement in
the diagonal Abelian effective theory; off-diagonal massive fields and
non-Abelian core effects are outside the hypotheses.

At the Chebyshev-type maximally singular point of the pure
\(\mathcal N=2\) \(SU(N_c)\) Coulomb branch, the Abelianized low-energy
calculation often produces an FI profile proportional to the positive
\(A_{N_c-1}\) sine vector
\begin{equation}
  s_i=\sin\frac{\pi i}{N_c},\qquad i=1,\ldots,N_c-1.
  \label{eq:an-sine-vector}
\end{equation}
The finite algebra behind the resulting sine ratios is elementary and is
recorded here because it is the part of the comparison that is independent of
any literature convention.

\begin{example}[\(A_{N_c-1}\) sine profile and normalized tensions]
\label{ex:a-type-sine-profile-k-string}
Let \(s_i=\sin(\pi i/N_c)\).  Then \(s=(s_i)\) is the strictly positive
eigenvector of the \(A_{N_c-1}\) Cartan matrix with eigenvalue
\begin{equation}
  \lambda_1=4\sin^2\frac{\pi}{2N_c}.
  \label{eq:a-type-sine-profile-eigenvalue}
\end{equation}
Consequently, if an Abelianized \(SU(N_c)\) string construction has aligned FI
magnitudes
\begin{equation}
  |\xi_k|=|\xi_1|\frac{\sin(\pi k/N_c)}{\sin(\pi/N_c)}
  \qquad (1\le k\le N_c-1),
  \label{eq:sw-sine-fi-profile}
\end{equation}
and if the \(k\)-labelled string is the single BPS sector associated with
\(\xi_k\), then the Abelianized normalized tensions are
\begin{equation}
  \frac{T_k^{\rm SW}}{T_1^{\rm SW}}
  =
  \frac{\sin(\pi k/N_c)}{\sin(\pi/N_c)}.
  \label{eq:sw-sine-k-string-ratio}
\end{equation}
They obey charge conjugation
\begin{equation}
  T_k^{\rm SW}=T_{N_c-k}^{\rm SW}
  \label{eq:sw-sine-charge-conjugation}
\end{equation}
and strict subadditivity
\begin{equation}
  T_{k+\ell}^{\rm SW}<T_k^{\rm SW}+T_\ell^{\rm SW},
  \qquad
  k,\ell\ge1,\quad k+\ell\le N_c-1 .
  \label{eq:sw-sine-subadditivity}
\end{equation}
\end{example}

The verification is finite-dimensional.  For \(1<i<N_c-1\),
\[
  (Ks)_i
  =
  2\sin\frac{\pi i}{N_c}
  -\sin\frac{\pi(i-1)}{N_c}
  -\sin\frac{\pi(i+1)}{N_c}.
\]
Using
\[
  \sin(x-a)+\sin(x+a)=2\sin x\cos a
\]
with \(a=\pi/N_c\), this becomes
\[
  2\left(1-\cos\frac{\pi}{N_c}\right)
  \sin\frac{\pi i}{N_c}
  =
  4\sin^2\frac{\pi}{2N_c}\,s_i .
\]
The endpoint rows obey the same identity with the absent sine terms
\(\sin0\) and \(\sin\pi\), both zero.  The entries \(s_i\) are positive for
\(1\le i\le N_c-1\), so this is the Perron--Frobenius eigenvector of the
Cartan matrix.

Under the FI profile \eqref{eq:sw-sine-fi-profile}, the saturated rank-one
BPS tension in the \(k\)-sector is \(T_k^{\rm SW}=2\pi|\xi_k|\).  Division by
the \(k=1\) result gives \eqref{eq:sw-sine-k-string-ratio}.  Charge
conjugation follows from
\[
  \sin\frac{\pi(N_c-k)}{N_c}
  =
  \sin\frac{\pi k}{N_c}.
\]
For subadditivity, set \(a=\pi k/N_c\) and \(b=\pi\ell/N_c\).  Since
\(0<a,b\) and \(a+b<\pi\),
\[
  \sin a+\sin b-\sin(a+b)
  =
  4\sin\frac a2\sin\frac b2\sin\frac{a+b}{2}>0 .
\]
After multiplying by the positive normalization
\((2\pi|\xi_1|)/\sin(\pi/N_c)\), this gives
\eqref{eq:sw-sine-subadditivity}.

\begin{definition}[Seiberg--Witten to pure-Yang--Mills string comparison]
\label{def:sw-to-pure-ym-string-comparison}
A comparison between the Abelianized Seiberg--Witten strings and the bosonic
pure-Yang--Mills \(k\)-strings consists of:
\begin{enumerate}[label=(\roman*)]
\item a fixed global form of the gauge group and a fixed line-operator
  lattice, so that the center charge \(k\) means the same thing throughout;
\item normalized tension functions
  \[
    R_k(t)=\frac{T_k(t)}{T_1(t)}
  \]
  along a declared deformation path \(t\), where \(t=0\) denotes the
  Abelianized Seiberg--Witten regime and \(t=1\) denotes the target bosonic
  Yang--Mills problem;
\item a statement of the topology in which the flux-sector transfer matrices,
  their lowest eigenvalues, and the large-length limits defining \(T_k(t)\)
  vary with \(t\);
\item a list of possible obstructions: loss of the Abelian patch, string
  breaking by dynamical matter, collision with another flux-sector level,
  change of line-operator lattice, or failure of the continuum limit.
\end{enumerate}
\end{definition}

\paragraph{What the Abelian sine calculation does and does not transfer.}
Assume a comparison problem as in
Definition~\ref{def:sw-to-pure-ym-string-comparison}.  If, for a fixed
\(k\), the tensions \(T_1(t)\) and \(T_k(t)\) exist, are positive, and are
absolutely continuous on an interval \(I\), then the only transferable
numerical information is the initial value together with a bound on the
relative response
\begin{equation}
  \Gamma_k(t)
  =
  \frac{\partial_tT_k(t)}{T_k(t)}
  -
  \frac{\partial_tT_1(t)}{T_1(t)} .
  \label{eq:sw-string-ratio-response}
\end{equation}
Indeed, for \(t_0,t_1\in I\),
\begin{equation}
  \log\frac{R_k(t_1)}{R_k(t_0)}
  =
  \int_{t_0}^{t_1}\Gamma_k(t)\,\dd t,
  \qquad
  \left|\log\frac{R_k(t_1)}{R_k(t_0)}\right|
  \le
  \int_{t_0}^{t_1}|\Gamma_k(t)|\,\dd t .
  \label{eq:sw-string-ratio-transport-bound}
\end{equation}
The Abelian Seiberg--Witten calculation proves the initial value
\(R_k(0)\) under its low-energy hypotheses.  It does not by itself bound
\(\Gamma_k\) after the path leaves the Abelian patch.  Setting
\(\Gamma_k=0\), or assuming the integral in
\eqref{eq:sw-string-ratio-transport-bound} vanishes, is precisely the
nonperturbative flux-sector statement that must be proved from a common
regulator, transfer matrices, level isolation, and continuum-limit control.
Thus the sine ratio is a controlled Abelian initial datum; its equality to a
bosonic pure-Yang--Mills ratio is not asserted in this chapter.

\subsection[Planar N=4]{Planar \(\mathcal N=4\)}

Planar \(\mathcal N=4\) Yang--Mills is conformal, so the family comparison uses
scaling dimensions and defect energies instead of massive glueball poles.  A
local-operator row is the planar spin-chain/QSC spectral problem of
Chapters~\ref{ch:planar-n4-spectral-problem-spin-chains}--\ref{ch:planar-n4-quantum-spectral-curve-hexagon}.  A line-defect row is the
Wilson-line flux-tube spectral problem of
Definition~\ref{def:planar-n4-wilson-line-flux-tube-problem}.  Its vacuum
energy is the cusp anomalous dimension, and its excitations are the
pentagon-OPE or mirror-TBA flux-tube excitations in an integrability
description.

The \(\mathcal N=4\) row is therefore the most exactly soluble spectral row
currently available inside four-dimensional interacting QFT, while its
proof boundary is also sharp: the planar integrability construction is an
exact spectral framework under stated integrability hypotheses, and a
derivation from a nonperturbative four-dimensional regulator remains open.

\section{Soft Breaking Toward Bosonic Yang--Mills}
\label{sec:susy-ym-family-soft-breaking-to-bosonic}

The deformation from pure \(\mathcal N=1\) Yang--Mills toward bosonic
Yang--Mills is naturally expressed by a spurion source for the chiral
coordinate \(S\):
\begin{equation}
  \Delta\mathcal L
  =
  \int\dd^2\theta\,X_\lambda(\theta)S+\mathrm{c.c.},
  \qquad
  X_\lambda(\theta)=\theta^2m_\lambda .
  \label{eq:gaugino-mass-spurion-source}
\end{equation}
The component deformation is
\begin{equation}
  \Delta\mathcal L_{\rm comp}
  =
  m_\lambda\,S|_{\theta=0}
  +
  \overline m_\lambda\,\overline S|_{\overline\theta=0}.
  \label{eq:gaugino-mass-component-source}
\end{equation}
At small \(|m_\lambda|\), the leading branch energy splitting is obtained
from the chiral condensate:
\begin{equation}
  \Delta E_k
  =
  -2\,\operatorname{Re}\!\left(
    m_\lambda\,C\Lambda_h^3\ee^{2\pi\ii k/N_c}
  \right)
  +O(|m_\lambda|^2).
  \label{eq:gaugino-mass-branch-splitting}
\end{equation}
This formula is a first-order source-response statement in the massive
\(\mathcal N=1\) theory.  Increasing \(|m_\lambda|\) until the gaugino
decouples is a deformation path whose spectral continuity must be examined
with nonchiral observables: glueball masses, string tensions, domain-wall
fates, and line-operator sectors.  The endpoint comparison with bosonic
Yang--Mills therefore belongs to the spectral comparison table of
Definition~\ref{def:susy-ym-family-observable-table}, not only to the
chiral branch comparison.

\begin{controlledapproximation}[Small soft mass branch selection]
\label{ca:small-soft-mass-branch-selection}
Assume pure \(\mathcal N=1\) \(SU(N_c)\) Yang--Mills has \(N_c\) massive
branches with condensates
\[
  S_k=C\Lambda_h^3\ee^{2\pi\ii k/N_c}
\]
and that the soft source \eqref{eq:gaugino-mass-spurion-source} can be
treated in first-order perturbation theory in a finite volume before the
thermodynamic limit is taken.  Then the branch selected at first order is the
branch maximizing
\[
  \operatorname{Re}(m_\lambda S_k).
\]
\end{controlledapproximation}

\paragraph{First-order branch selection.}
The first-order shift of the vacuum energy density under the component source
\(m_\lambda S+\overline m_\lambda\overline S\) is minus the expectation value
of the source term:
\[
  \Delta E_k
  =
  -
  \left(m_\lambda S_k+\overline m_\lambda\,\overline S_k\right)
  +O(|m_\lambda|^2)
  =
  -2\operatorname{Re}(m_\lambda S_k)+O(|m_\lambda|^2).
\]
At first order, minimizing \(\Delta E_k\) is the same as maximizing
\(\operatorname{Re}(m_\lambda S_k)\).

\subsection{A Controlled Soft-Mass Spectral Bridge Segment}
\label{sec:susy-soft-mass-spectral-bridge-segment}

The small-\(m_\lambda\) deformation has one place where the chapter can make
a genuine spectral bridge rather than a comparison slogan: finite-volume
spectral projections may be transported as long as the relevant levels remain
isolated.  This is a local bridge in the soft-mass parameter.  It is not a
proof that the path reaches bosonic Yang--Mills.

\begin{definition}[Common regulated soft-mass spectral bridge datum]
\label{def:soft-mass-spectral-bridge-datum}
A regulated soft-mass spectral bridge datum consists of:
\begin{enumerate}[label=(\roman*)]
\item a common gauge-field regulator with lattice spacing or cutoff \(a\), a
  finite spatial box of size \(L\), a fixed global form, and a fixed
  line-operator lattice;
\item for each sector \(\sigma\) used below, a Hilbert space
  \(\mathcal H_{a,L}^{\sigma}\) and a self-adjoint Hamiltonian
  \(H_{a,L}^{\sigma}(m)\).  In a compact Hamiltonian lattice regulator this
  means the \(L^2\)-space of link variables, tensored with the finite fermion
  Fock space and projected to the Gauss-law and flux sector.  The
  electric-field kinetic term is unbounded, so the sector comes with a dense
  self-adjoint domain \(\mathcal D_{\sigma}\), or with a closed quadratic-form
  domain \(\mathcal Q_{\sigma}\);
\item a renormalized one-parameter path \(H_{a,L}^{\sigma}(m)\) along a fixed
  ray \(m=\operatorname{Re}m_\lambda\) in the soft source plane, including the
  counterterm and tuning prescriptions that keep the gauge coupling, vacuum
  energy, and relevant fermion-mass coordinate in the declared scheme;
\item an operator-topology hypothesis for that path.  For the operator
  formulae below it is enough to assume a common-domain Kato type-\((A)\)
  family: \(\mathcal D_{\sigma}\) is independent of \(m\),
  \(m\mapsto H_{a,L}^{\sigma}(m)\psi\) is \(C^1\) for every
  \(\psi\in\mathcal D_{\sigma}\), and the derivative is relatively bounded
  on every compact \(m\)-interval,
  \begin{equation}
    \|\mathcal V_{a,L}^{\sigma}(m)\psi\|
    \le
    A_{a,L,I}\|H_{a,L}^{\sigma}(m)\psi\|
    +B_{a,L,I}\|\psi\|,
    \qquad \psi\in\mathcal D_{\sigma}.
    \label{eq:soft-mass-relative-bound}
  \end{equation}
  A closed-quadratic-form type-\((B)\) hypothesis may also be used, with a
  common form domain \(\mathcal Q_{\sigma}\) and a \(C^1\) form derivative
  bounded by the Hamiltonian form plus \(\|\cdot\|^2\).  In that case the
  displayed matrix elements of \(\mathcal V\) below are read as form
  derivatives.  A direct \(C^1\) norm-resolvent hypothesis is an equally
  sufficient substitute;
\item the renormalized perturbing operator or form derivative
  \begin{equation}
    \mathcal V_{a,L}^{\sigma}(m)=\partial_m H_{a,L}^{\sigma}(m),
    \label{eq:soft-mass-perturbing-operator}
  \end{equation}
  whose bare representative is the spatial integral of
  \(S|_{\theta=0}+\overline S|_{\overline\theta=0}\) plus the required
  contact and counterterm derivatives, all satisfying the same relative or
  form bounds as the complete derivative;
\item isolated finite-volume spectral windows for a selected vacuum branch,
  for gauge-invariant local channels, and for selected center-flux sectors;
\item explicit gap, no-level-collision, and resolvent bounds on those windows
  for \(0\le |m|\le m_\ast(a,L)\), together with the requirement that every
  finite-rank Riesz range used in a trace formula lies in
  \(\mathcal D_{\sigma}\), or in \(\mathcal Q_{\sigma}\) for the form
  version.
\end{enumerate}
\end{definition}

\begin{controlledapproximation}[Finite-volume soft-mass spectral transport]
\label{ca:soft-mass-finite-volume-spectral-transport}
Assume the datum of
Definition~\ref{def:soft-mass-spectral-bridge-datum} in a fixed sector
\(\sigma\), and use the type-\((A)\) or \(C^1\) norm-resolvent hypothesis
for the displayed operator formulae.  Let \(I_n(0)\) be an isolated
local-channel spectral cluster of \(H_{a,L}^{\sigma}(0)\), separated from the
rest of the finite-volume spectrum by distance \(g_n>0\), and let
\(\mathcal C_n\) be a contour enclosing only that cluster.  Suppose
\(\mathcal C_n\) remains in the resolvent set of \(H_{a,L}^{\sigma}(m)\)
throughout the interval and
\[
  \sup_{z\in\mathcal C_n}
  \|(z-H_{a,L}^{\sigma}(m))^{-1}
    \mathcal V_{a,L}^{\sigma}(m)
    (z-H_{a,L}^{\sigma}(m))^{-1}\|
  <\infty .
\]
Then the Riesz projection
\begin{equation}
  P_n(m)=\frac{1}{2\pi\ii}
  \oint_{\mathcal C_n}(z-H_{a,L}^{\sigma}(m))^{-1}\,\dd z
  \label{eq:soft-mass-riesz-projection}
\end{equation}
exists with constant rank and satisfies
\begin{equation}
  \partial_mP_n(m)
  =
  \frac{1}{2\pi\ii}
  \oint_{\mathcal C_n}
  (z-H_{a,L}^{\sigma}(m))^{-1}\mathcal V_{a,L}^{\sigma}(m)
  (z-H_{a,L}^{\sigma}(m))^{-1}\,\dd z .
  \label{eq:soft-mass-projection-derivative}
\end{equation}
For the cluster centroid
\[
  \overline E_n(m)
  =
  \frac{1}{d_n}
  \operatorname{Tr}\bigl(P_n(m)H_{a,L}^{\sigma}(m)\bigr),
  \qquad d_n=\operatorname{rank}P_n,
\]
the mass above the selected vacuum obeys the finite-volume
Feynman--Hellmann identity
\begin{equation}
  \partial_m\bigl(\overline E_n(m)-E_0(m)\bigr)
  =
  \frac{1}{d_n}
  \operatorname{Tr}\bigl(P_n(m)\mathcal V_{a,L}^{\sigma}(m)P_n(m)\bigr)
  -
  \langle\Omega_m,\mathcal V_{a,L}^{0}(m)\Omega_m\rangle .
  \label{eq:soft-mass-cluster-feynman-hellmann}
\end{equation}
For a nondegenerate level this reduces to the difference of the perturbing
operator's matrix element in the excited state and in the vacuum.  The same
statement holds in a center-flux sector \(k\), with the corresponding
finite-volume lowest energy \(E_k(L,m)\):
\begin{equation}
  \partial_m T_k^{(L)}(m)
  =
  \frac1L\left(
    \langle k,L;m|\mathcal V_{a,L}^{k}(m)|k,L;m\rangle
    -
    \langle\Omega_m,\mathcal V_{a,L}^{0}(m)\Omega_m\rangle
  \right),
  \label{eq:soft-mass-flux-feynman-hellmann}
\end{equation}
where \(T_k^{(L)}(m)=(E_k(L,m)-E_0(L,m))/L\).  If the
\(a\to0\), \(L\to\infty\), and \(m\)-derivative limits are uniform in the
declared order, then \eqref{eq:soft-mass-flux-feynman-hellmann} gives the
continuum string-tension response.  In that case
\begin{equation}
  \partial_m\log R_k(m)
  =
  \frac{\partial_mT_k(m)}{T_k(m)}
  -
  \frac{\partial_mT_1(m)}{T_1(m)}
  \label{eq:soft-mass-ratio-response}
\end{equation}
is the quantity that controls whether the normalized \(k\)-string ratio is
transported.
\end{controlledapproximation}

\paragraph{Operator-theoretic proof of the local bridge.}
The common-domain type-\((A)\) or norm-resolvent hypothesis supplies the
resolvent derivative in operator norm,
\begin{equation}
  \partial_m(z-H_{a,L}(m))^{-1}
  =
  (z-H_{a,L}(m))^{-1}\mathcal V_{a,L}(m)(z-H_{a,L}(m))^{-1},
  \label{eq:soft-mass-resolvent-derivative-domain}
\end{equation}
with sector labels suppressed.  For a type-\((A)\) family this follows by
differentiating the common-domain identity
\[
  (z-H_{a,L}(m))R_z(m)=1,
  \qquad
  R_z(m):=(z-H_{a,L}(m))^{-1},
\]
and extending the result by the relative bound
\eqref{eq:soft-mass-relative-bound}; it is the closed-operator perturbation
theorem used in Kato--Reed--Simon perturbation theory
\cite{Kato1966Perturbation,ReedSimon1975II}.  A type-\((B)\) datum
gives the same statement in form-resolvent language and replaces
\(\langle\psi,\mathcal V\psi\rangle\) by the derivative of the closed form on
\(\mathcal Q_{\sigma}\).

Integrating \eqref{eq:soft-mass-resolvent-derivative-domain} around
\(\mathcal C_n\) gives
\eqref{eq:soft-mass-projection-derivative}.  Norm-continuity of the Riesz
projection and absence of contour crossings keep the rank constant.  The
resolvent maps into the operator domain, so the finite-rank range of
\(P_n(m)\) lies in the domain required for the trace against
\(\mathcal V_{a,L}^{\sigma}(m)\).  Differentiating
\(\operatorname{Tr}(P_nH_{a,L}^{\sigma})\) gives two terms:
\[
  \operatorname{Tr}((\partial_mP_n)H_{a,L}^{\sigma})
  +
  \operatorname{Tr}(P_n\mathcal V_{a,L}^{\sigma}).
\]
The first vanishes because \(P_n\) commutes with \(H_{a,L}^{\sigma}\) and
\(\operatorname{Tr}(\partial_mP_n\,P_n)=0\) for a differentiable projection of
fixed rank.  Subtracting the vacuum derivative gives
\eqref{eq:soft-mass-cluster-feynman-hellmann}.  The flux-sector formula is
the same identity applied to the transfer-matrix Hamiltonian in the sector
with center charge \(k\), followed by division by the long length \(L\).

This bridge has concrete failure diagnostics.  It fails before the bosonic
endpoint if any selected local or flux spectral gap closes, if the Riesz
contour intersects another level, if the common operator domain or common
form domain is lost, if the relative or form bound for the soft-mass
insertion and counterterm derivatives blows up, if center flux can break by
dynamical matter in the chosen global form, if the fermion regulator changes
Pfaffian or reflection-positivity status, or if the counterterm/tuning
trajectory needed to define \(\mathcal V_{a,L}\) is not uniform in the
continuum and thermodynamic limits.  These are the phase-boundary tests for
the soft-mass path.

\paragraph{Finite cutoff versus continuum endpoint.}
All hypotheses in
Definition~\ref{def:soft-mass-spectral-bridge-datum} are finite-cutoff
operator hypotheses at fixed \(a,L\).  They justify
\eqref{eq:soft-mass-cluster-feynman-hellmann} and
\eqref{eq:soft-mass-flux-feynman-hellmann} inside the regulated Hilbert
spaces.  The passage to \(a\to0\), \(L\to\infty\), and then to a large
soft-gaugino-mass decoupling regime requires additional uniform spectral
gaps, uniform relative bounds, reflection-positive transfer matrices or an
equivalent Hamiltonian construction, and phase-boundary exclusion.  The local
bridge supplies a controlled derivative and stability criterion while those
finite-cutoff tests pass; the bosonic Yang--Mills endpoint remains a
separate theorem boundary.

\section{Status Summary}
\label{sec:susy-ym-family-status-summary}

The present chapter closes the elementary comparison map.  It records the
following theorem boundaries:
\[
\begin{array}{c|c}
\text{statement} & \text{status in this monograph}\\
\hline
\mathcal N=4\to\mathcal N=2,\mathcal N=1
\text{ scale matching} & \text{derived holomorphic threshold matching}\\
\mathcal N=1\text{ branches and BPS walls} &
  \text{central conjecture plus chiral }F\text{-term evidence}\\
\mathcal N=2\text{ monopole condensation} &
  \text{derived in local Abelian patch}\\
\mathcal N=2\text{ vortex tension} &
  \text{BPS Abelian-Higgs controlled approximation}\\
\mathcal N=2\text{ Abelianized }k\text{-strings} &
  \text{\(A\)-type sine profile; comparison problem}\\
\text{small }m_\lambda\text{ spectral transport} &
  \text{finite-volume FH response; isolated windows}\\
\mathcal N=4\text{ Wilson-line flux tube} &
  \text{defect spectral problem plus integrability hypotheses}\\
\text{bosonic glueball and }k\text{-string spectra} &
  \text{central YM conjecture plus spectra}
\end{array}
\]
The frontier problem is to extend the local soft-mass bridge of
Controlled Approximation~\ref{ca:soft-mass-finite-volume-spectral-transport}
to a regulator-level
family in which enough of the ladder is realized simultaneously to compare the
spectral measures, line-defect Hilbert spaces, and flux-sector transfer
matrices all the way to a bosonic endpoint.

\begin{openproblem}[Regulator-level supersymmetry-breaking spectral bridge]
\label{op:regulator-level-susy-breaking-spectral-bridge}
Construct a regulator and continuum-limiting framework for the deformation
from supersymmetric Yang--Mills theories to bosonic Yang--Mills that controls
all of the following in the same topology: gauge-invariant local spectral
measures, center-flux transfer matrices, line-operator charge lattices,
domain-wall sectors, and protected chiral coordinates.  Prove, disprove, or
replace the commonly assumed continuity of glueball and flux-string spectra
along such deformation paths by a theorem with stated hypotheses.
\end{openproblem}
