\chapter{Supersymmetric Wilsonian Schemes and Exact Dynamics}
\label{ch:susy-wilsonian-schemes-exact-dynamics}

\ConventionsLorentzian

Exact statements in supersymmetric QFT require a declared Wilsonian scheme.
The scheme is the data that define which modes are integrated out, which
local coordinates are used for the resulting action, which Ward identities
hold at finite cutoff, and which anomaly classes obstruct stronger
statements.  Supersymmetry of the Wilsonian integration and supersymmetry of
the regulator are separate requirements.

\section{Wilsonian Data}

\begin{definition}[Supersymmetric Wilsonian datum]
\label{def:susy-wilsonian-datum}
A supersymmetric Wilsonian datum at scale \(\Lambda\) consists of:
a space \(\mathcal E_\Lambda\) of regulated superfield variables, a gauge
group action together with either a BV gauge complex or a lattice
gauge-field construction, a projection
\(\pi_{\Lambda\Lambda'}:\mathcal E_{\Lambda}\to\mathcal E_{\Lambda'}\) for
\(\Lambda>\Lambda'\), an integration cycle in the discarded variables, and a
local action functional \(S_\Lambda\) satisfying the declared gauge and
supersymmetry Ward identities at finite cutoff.
\end{definition}

If the gauge symmetry is present, the natural finite-cutoff statement is a BV
statement.  Let \((\,\cdot\,,\,\cdot\,)_\Lambda\) be the odd BV bracket and
\(\Delta_\Lambda\) the regulated BV Laplacian.  The quantum master equation is
\[
  \frac12(S_\Lambda,S_\Lambda)_\Lambda
  -\ii\hbar\,\Delta_\Lambda S_\Lambda=0 .
\]
This equation is meaningful only after the regulator and the functional
measure entering \(\Delta_\Lambda\) have been specified.
For a gauge-fixed supersymmetric Wilsonian path integral, gauge consistency is
the same equation before restriction to the gauge-fixing Lagrangian
submanifold.  The gauge-fixed Slavnov--Taylor identities and supersymmetry
Ward identities are obtained by restricting the BV half-density and applying
BV Stokes.  A Wilsonian coarse-graining step is gauge-consistent when it is a
BV pushforward that sends a solution of the quantum master equation at
\(\Lambda\) to a solution at \(\Lambda'\).

A lattice supersymmetric gauge construction, when it exists, is the other
finite-cutoff route: gauge fields are link variables, gauge invariance is
exact at finite lattice spacing, and Wilsonian coarse graining is a blocking
map for link variables or gauge-invariant observables.  The obstruction is
then shifted to the realization of supersymmetry, fermions, auxiliary
variables, reflection positivity, and the continuum limit.  The two routes are
therefore complementary: BV is the continuum homological framework for
gauge-compatible effective actions, while the lattice gives a compact
gauge-invariant finite integral.

\section{Coarse Graining as BV Pushforward}

For \(\Lambda>\Lambda'\), a Wilsonian step must be defined before one writes
an action \(S_{\Lambda'}\).  The object naturally transported by BV
integration is not the action itself but the BV half-density
\[
  \rho_\Lambda
  =
  \exp\!\left(\frac{\ii}{\hbar}S_\Lambda\right)\mu_\Lambda^{1/2},
\]
where \(\mu_\Lambda^{1/2}\) is the chosen regulated half-density.  The
quantum master equation is
\[
  \Delta_\Lambda\rho_\Lambda=0,
\]
equivalently the displayed equation for \(S_\Lambda\) when
\(\Delta_\Lambda\) is written using \(\mu_\Lambda^{1/2}\).

\begin{definition}[Admissible BV Wilsonian step]
\label{def:admissible-bv-wilsonian-step}
An admissible BV Wilsonian step from \(\Lambda\) to \(\Lambda'\) consists of
the following finite-cutoff data.

\begin{enumerate}[label=(\roman*)]
\item A split exact sequence of odd symplectic spaces
\[
  0\longrightarrow K_{\Lambda\Lambda'}
  \longrightarrow \mathcal E_\Lambda
  \overset{p_{\Lambda\Lambda'}}{\longrightarrow}
  \mathcal E_{\Lambda'}
  \longrightarrow 0
\]
such that, after choosing the splitting,
\[
  \omega_\Lambda=\omega_{\Lambda'}\oplus\omega_{K_{\Lambda\Lambda'}},
  \qquad
  \Delta_\Lambda=\Delta_{\Lambda'}+\Delta_{K_{\Lambda\Lambda'}} .
\]
\item A Lagrangian integration cycle
\(\mathcal L_{\Lambda\Lambda'}\subset K_{\Lambda\Lambda'}\) for the odd
symplectic form on the eliminated variables, together with an orientation
or Berezinian convention fixing the fiber integral.
\item A convergence or oscillatory-integral prescription for which the fiber
integral of the half-density is defined and for which differentiation by
\(\Delta_{\Lambda'}\) may be interchanged with the fiber integral.
\item A BV boundary condition:
\begin{equation}
  \int_{\mathcal L_{\Lambda\Lambda'}}
  \Delta_{K_{\Lambda\Lambda'}}\alpha=0
  \label{eq:admissible-bv-boundary-condition}
\end{equation}
for every half-density \(\alpha\) in the class generated by the Wilsonian
integrand and by the variations used in Ward identities.
\item If supersymmetry is required at the cutoff, the regulated
supersymmetry generators preserve the split BV structure and the cycle
up to a deformation whose boundary term satisfies
\eqref{eq:admissible-bv-boundary-condition}.
\end{enumerate}
\end{definition}

The adjective ``admissible'' is load-bearing.  Item (i) supplies a genuine BV
factorization of fields, ghosts, antifields, and possible auxiliary variables;
a projection of bosonic gauge potentials alone is only a map of coordinates.
Item (ii) chooses the integration cycle for the eliminated variables, and
item (iii) supplies the analytic meaning of the fiber integral.  Item (iv) is
the finite-cutoff form of the statement that integrating a BV divergence over
the eliminated variables produces no boundary contribution.  In perturbative
language this condition is where the regulator, gauge fixing, contour, and
large-field behavior of the effective integration cycle enter.  Supersymmetry
at the cutoff is an extra compatibility datum in item (v): the regulated
supercharges must act on the full BV complex and preserve the admissible cycle
up to a BV-exact deformation with vanishing boundary term.  The proposition
below is therefore a BV Stokes consequence of these data; the construction of
such data is the substantive problem in a supersymmetric Wilsonian scheme.

\begin{lemma}[BV Stokes in a finite-cutoff Darboux chart]
\label{lem:finite-cutoff-bv-stokes}
Let \(K\) be a finite-dimensional odd symplectic space with Darboux
coordinates \(x^i,\xi_i\) and BV Laplacian
\[
  \Delta_K=\sum_i\frac{\partial^2}{\partial x^i\partial \xi_i}
\]
on functions multiplying the standard half-density.  If the Lagrangian cycle
is locally \(\xi_i=0\), then
\[
  \int_{\xi=0}\Delta_K\alpha
  =
  \int_{\partial(\xi=0)}
  \sum_i
  \left.
  \frac{\partial \alpha}{\partial \xi_i}
  \right|_{\xi=0}
  \iota_{\partial/\partial x^i}(dx^1\cdots dx^n).
\]
In particular, the integral vanishes for compact boundaryless cycles and for
noncompact cycles on which the boundary term at infinity is zero.  The same
statement in another Darboux chart follows by the invariance of the BV
Laplacian on half-densities under canonical coordinate changes.
\end{lemma}

\begin{proof}
On \(\xi=0\),
\[
  \left.\Delta_K\alpha\right|_{\xi=0}
  =
  \sum_i
  \frac{\partial}{\partial x^i}
  \left.
  \frac{\partial\alpha}{\partial\xi_i}
  \right|_{\xi=0}.
\]
Integration over the \(x\)-cycle converts this ordinary divergence into the
displayed boundary integral.  The vanishing criterion is therefore the usual
Stokes theorem applied to the integration cycle.  Since half-density BV
integration is defined invariantly, a general Lagrangian chart gives the
same formula after a canonical change of Darboux coordinates.
\end{proof}

The Wilsonian pushforward associated to these data is
\begin{equation}
  (p_{\Lambda\Lambda'})_*\rho_\Lambda(\phi')
  =
  \int_{\mathcal L_{\Lambda\Lambda'}}
  \rho_\Lambda(\phi',\chi).
  \label{eq:bv-wilsonian-pushforward-half-density}
\end{equation}
Only after this half-density has been computed and a nonvanishing reference
half-density \(\mu_{\Lambda'}^{1/2}\) has been chosen do we write
\[
  (p_{\Lambda\Lambda'})_*\rho_\Lambda
  =
  \exp\!\left(\frac{\ii}{\hbar}S_{\Lambda'}\right)
  \mu_{\Lambda'}^{1/2}.
\]
The logarithm is therefore a coordinate choice on half-densities; the
semigroup law is first a statement about pushforwards.

\begin{proposition}[BV pushforward preserves the quantum master equation]
\label{prop:bv-pushforward-preserves-qme}
For an admissible BV Wilsonian step,
\[
  \Delta_\Lambda\rho_\Lambda=0
  \qquad\Longrightarrow\qquad
  \Delta_{\Lambda'}(p_{\Lambda\Lambda'})_*\rho_\Lambda=0 .
\]
\end{proposition}

\begin{proof}
Using the split BV Laplacian,
\[
  \Delta_{\Lambda'}(p_{\Lambda\Lambda'})_*\rho_\Lambda
  =
  \int_{\mathcal L_{\Lambda\Lambda'}}
  \Delta_{\Lambda'}\rho_\Lambda
  =
  \int_{\mathcal L_{\Lambda\Lambda'}}
  \left(\Delta_\Lambda-\Delta_{K_{\Lambda\Lambda'}}\right)\rho_\Lambda .
\]
The first term vanishes by the quantum master equation at scale
\(\Lambda\).  The second term vanishes by the BV boundary condition
\eqref{eq:admissible-bv-boundary-condition}.  This is the finite-cutoff form
of BV Stokes: a BV Laplacian in the integrated variables is a total
divergence on the Lagrangian cycle, and the declared class of cycles has no
boundary contribution.
\end{proof}

For gauge theories, this BV compatibility or a lattice blocking construction
is part of the definition of the Wilsonian step.  A projection of gauge
potentials by itself does not define a gauge-theory coarse graining; the
projection must be lifted to the full ghost-antifield complex with a
compatible half-density and integration cycle, or replaced by a finite
gauge-invariant lattice blocking map.  When a gauge fixing or a
supersymmetric regulator changes with \(\Lambda\), the change is harmless
only if it is represented by an admissible BV isotopy of cycles; otherwise
the failure of \eqref{eq:admissible-bv-boundary-condition} is precisely a
Ward-identity anomaly of the scheme.

\begin{definition}[Composable BV Wilsonian tower]
\label{def:composable-bv-wilsonian-tower}
For \(\Lambda>\Lambda'>\Lambda''\), an admissible tower is composable if the
chosen splittings identify
\[
  \mathcal E_\Lambda
  \simeq
  \mathcal E_{\Lambda''}
  \oplus K_{\Lambda'\Lambda''}
  \oplus K_{\Lambda\Lambda'}
\]
as odd symplectic spaces, if
\[
  \mathcal L_{\Lambda\Lambda''}
  =
  \mathcal L_{\Lambda'\Lambda''}
  \times
  \mathcal L_{\Lambda\Lambda'}
\]
with the product Berezinian convention, and if the corresponding ordinary
or oscillatory Fubini theorem holds for the Wilsonian half-density and its
BV variations.
\end{definition}

\begin{proposition}[Semigroup property of BV Wilsonian pushforward]
\label{prop:susy-wilsonian-semigroup}
For a composable BV Wilsonian tower,
\[
  (p_{\Lambda'\Lambda''})_*
  (p_{\Lambda\Lambda'})_*\rho_\Lambda
  =
  (p_{\Lambda\Lambda''})_*\rho_\Lambda .
\]
Consequently, in any action coordinate for which the three half-densities are
written as
\(\rho_\bullet=\exp(\ii S_\bullet/\hbar)\mu_\bullet^{1/2}\), the Wilsonian
maps obey
\[
  S_{\Lambda''}
  =
  \mathcal R_{\Lambda'\Lambda''}
  \bigl(\mathcal R_{\Lambda\Lambda'}(S_\Lambda)\bigr)
  =
  \mathcal R_{\Lambda\Lambda''}(S_\Lambda)
\]
up to the additive normalization constants fixed by the chosen reference
half-densities.
\end{proposition}

\begin{proof}
For a test point \(\phi''\in\mathcal E_{\Lambda''}\), the iterated
pushforward is
\[
  \int_{\mathcal L_{\Lambda'\Lambda''}}
  \left(
  \int_{\mathcal L_{\Lambda\Lambda'}}
  \rho_\Lambda(\phi'',\eta,\chi)
  \right),
\]
where \(\eta\in K_{\Lambda'\Lambda''}\) and
\(\chi\in K_{\Lambda\Lambda'}\).  By the Fubini part of
Definition~\ref{def:composable-bv-wilsonian-tower}, this equals
\[
  \int_{\mathcal L_{\Lambda'\Lambda''}\times
        \mathcal L_{\Lambda\Lambda'}}
  \rho_\Lambda(\phi'',\eta,\chi).
\]
The product cycle is, by definition, the direct cycle
\(\mathcal L_{\Lambda\Lambda''}\).  This is exactly the direct pushforward.
The equality of actions follows only after passing from half-densities to the
chosen action coordinates; any field-independent determinant factor is an
additive constant in \(S\) and is fixed by the normalization convention.
\end{proof}

\begin{proposition}[Finite circle-Darboux model of BV pushforward]
\label{prop:finite-circle-darboux-bv-pushforward}
Let the base variables be a periodic even coordinate \(y\) and its odd BV
partner \(\eta\), and let the eliminated variables be a periodic even
coordinate \(x\) and its odd BV partner \(\xi\).  Work on the finite algebra
generated by finitely many Fourier modes
\[
  e^{imy}e^{inx},\qquad m,n\in\mathbb Z,
\]
and by the odd variables \(\eta,\xi\).  Define
\[
  \Delta_B=\frac{\partial}{\partial y}\frac{\partial^L}{\partial\eta},
  \qquad
  \Delta_K=\frac{\partial}{\partial x}\frac{\partial^L}{\partial\xi},
  \qquad
  \Delta=\Delta_B+\Delta_K .
\]
Let the fiber pushforward be the normalized circle integral over the
Lagrangian cycle \(\xi=0\):
\[
  P(\alpha)(y,\eta)
  =
  \int_{S^1_x}\alpha(y,\eta,x,0)\,\frac{dx}{2\pi}.
\]
Then
\begin{equation}
  P(\Delta_K\alpha)=0,\qquad
  P(\Delta\alpha)=\Delta_B P(\alpha).
  \label{eq:finite-circle-bv-chain-map}
\end{equation}
Consequently, if \(\Delta\alpha=0\), then \(\Delta_B P(\alpha)=0\).  If two
eliminated circle-Darboux pairs are present and the product cycle is used,
then the corresponding pushforwards commute and equal the direct product
pushforward.
\end{proposition}

\begin{proof}
It is enough to check a monomial
\[
  \alpha
  =
  c\,e^{imy}e^{inx}\eta^a\xi^b,
  \qquad a,b\in\{0,1\}.
\]
If \(b=0\), then \(\partial^L\alpha/\partial\xi=0\), so
\(P(\Delta_K\alpha)=0\).  If \(b=1\), then
\(\Delta_K\alpha\) is \(n\) times a Fourier mode \(e^{inx}\) with no
\(\xi\)-factor.  The normalized circle integral of a nonzero derivative mode
vanishes, and the \(n=0\) mode has derivative coefficient \(n=0\).  Thus
\(P(\Delta_K\alpha)=0\) in all cases.  Since \(\Delta_B\) differentiates
only the base variables, it commutes with the normalized \(x\)-integral and
with setting \(\xi=0\).  This proves
\eqref{eq:finite-circle-bv-chain-map}.

For two eliminated pairs, each pushforward selects the zero Fourier mode in
its own even variable and the coefficient with no corresponding odd fiber
variable.  These two coefficient extractions act on disjoint variables, so
their order is irrelevant and the result is the same as extracting both at
once.  This is the finite Fourier model of the Fubini and product-cycle
assumption in Definition~\ref{def:composable-bv-wilsonian-tower}.
\end{proof}

\section{Local Coordinates: F-Terms and D-Terms}

For four-dimensional \(\mathcal N=1\) theories, local supersymmetric
functionals split by superspace measure:
\[
  \int d^4x\,d^2\theta\,\mathcal F(\Phi,W_\alpha,\nabla)
  +\int d^4x\,d^4\theta\,\mathcal K(\Phi,\bar\Phi,V,\nabla).
\]
The first class is called \(F\)-term coordinates and the second class
\(D\)-term coordinates.  This coordinate decomposition becomes a statement
about dynamics only after a Wilsonian scheme and its Ward identities have
been specified.  The gauge kinetic coordinate is the chiral functional
\[
  \frac{1}{16\pi\ii}\int d^4x\,d^2\theta\,
  \tau\,\operatorname{tr}(W^\alpha W_\alpha)+\text{c.c.},
  \qquad
  \tau=\frac{\theta}{2\pi}+\frac{4\pi\ii}{g^2},
\]
with the trace normalization fixed in
Chapter~\ref{ch:susy-gauge-theory}.  If a different superspace normalization
for \(W_\alpha\) is chosen, the displayed coefficient is changed by the
component-matching factor; the physical convention is the component
Yang--Mills term \(-\frac{1}{4g^2}\operatorname{tr}F^2\).

\section{Holomorphic Wilsonian Claims}

\begin{hypothesis}[Holomorphic Wilsonian scheme]
\label{hyp:holomorphic-wilsonian-scheme}
A Wilsonian scheme is holomorphic for a set of chiral couplings \(t^A\) if
the regulated chiral action depends holomorphically on \(t^A\), the
coarse-graining map preserves the chiral superspace cohomology class of local
\(F\)-terms, and all violations of holomorphy are represented by declared
local anomaly functionals.
\end{hypothesis}

Under this hypothesis, nonrenormalization statements become statements about
coordinates.  A superpotential coordinate
\[
  \int d^4x\,d^2\theta\, W_{\rm tree}(\Phi;t)
\]
can receive only those Wilsonian corrections that are holomorphic functions
of the chiral coordinates, respect the exact symmetries including anomalous
spurion transformations, and are local at scale \(\Lambda\).  The
selection-rule arguments are replaced here by these three stated conditions.

\paragraph{Selection rule for a Wilsonian \(F\)-term.}
Assume Hypothesis~\ref{hyp:holomorphic-wilsonian-scheme}.  Let
\(\mathcal U(t,\Phi)\) be a candidate correction to a chiral \(F\)-term.  If
\(\mathcal U\) is not a holomorphic section of the same symmetry line bundle
as the original chiral measure and operator insertion, then it is absent from
the Wilsonian \(F\)-term coordinate in that scheme.
The Wilsonian \(F\)-term coordinate is, by hypothesis, a holomorphic element
of the chiral local cohomology with specified transformation under exact and
anomalous symmetries.  Local terms with different transformation law lie in a
different line bundle over the coupling space.  They cannot be added to the
coordinate without violating the Ward identity defining the scheme.  Thus
only holomorphic sections with the same transformation law are admissible.

\section{Exact Dynamics as a Scheme-Dependent Statement}

The chiral ring at scale \(\Lambda\) is the cohomology of local chiral
operators modulo \(\bar Q\)-exact operators and equations of motion in the
Wilsonian scheme.  A moduli space statement is an assertion about expectation
values of this ring in supersymmetric states together with the effective
action governing fluctuations.  A duality statement is an isomorphism between
two such Wilsonian data, including line operators, anomalies, contact terms,
and relevant deformations.

\begin{openproblem}[Wilsonian construction]
\label{op:susy-wilsonian-construction}
Construct a Wilsonian scheme for four-dimensional gauge theories that
preserves gauge invariance, realizes supersymmetry in BV form at finite
cutoff, states the anomaly functionals, and proves the holomorphic hypotheses
used in exact dynamics.  This is the infrastructure needed for proofs of
\(\mathcal N=1\) and \(\mathcal N=2\) gauge dynamics.
\end{openproblem}
