\chapter{Planar N=4 Supersymmetric Yang-Mills as a Spectral Problem}
\label{ch:planar-n4-spectral-problem-spin-chains}

\ConventionsLorentzian

The subject of this chapter is the spectrum of local gauge-invariant
operators in four-dimensional \(\mathcal N=4\) supersymmetric Yang--Mills
theory in the planar limit.  The same theory appears earlier in this volume
as a maximally supersymmetric conformal gauge theory.  Here it is treated as a
spectral problem: diagonalize the dilatation operator on the vector space of
renormalized local operators.

\section{SCFT Data Before Planarity}
\label{sec:n4-scft-data-before-planarity}

The planar spin chain is not the definition of the theory.  It is a useful
coordinate system on a particular large-\(N\) sector of a four-dimensional
superconformal gauge theory.  We first record the intrinsic quantum-field
theory data that will remain meaningful at finite \(N\), away from any
spin-chain limit.

\begin{definition}[N=4 Yang-Mills field content]
\label{def:n4-yang-mills-field-content}
Fix a compact gauge group \(G\), Lie algebra \(\mathfrak g\), and invariant
trace form \(\operatorname{tr}\) as in
Chapter~\ref{ch:susy-gauge-theory}.  The four-dimensional \(\mathcal N=4\)
Yang--Mills field variables consist of
\[
  A_\mu,\qquad
  \Phi^I\ (I=1,\ldots,6),\qquad
  \lambda^A_\alpha,\ \bar\lambda_{A\dot\alpha}\ (A=1,\ldots,4),
\]
all valued in \(\mathfrak g\).  The six real scalars transform in the vector
representation of \(SO(6)_R\), equivalently in the antisymmetric
representation of \(SU(4)_R\); the four Weyl fermions transform in the
fundamental of \(SU(4)_R\).  In the monograph coupling convention the bosonic
part of the Lorentzian action is
\[
\begin{aligned}
  S_{\rm bos}
  &=
  \frac{1}{g_{\rm YM}^2}
  \int d^4x\,\operatorname{tr}\!\left(
  -\frac14 F_{\mu\nu}F^{\mu\nu}
  -\frac12 D_\mu\Phi^I D^\mu\Phi^I
  +\frac14[\Phi^I,\Phi^J][\Phi^I,\Phi^J]
  \right)
  \\
  &\quad+
  \frac{\theta}{32\pi^2}
  \int d^4x\,
  \operatorname{tr}(F_{\mu\nu}\widetilde F^{\mu\nu}) .
\end{aligned}
\]
together with the unique Yukawa couplings compatible with the same
\(\mathcal N=4\) supersymmetry and \(SU(4)_R\) action.  The complexified
gauge coupling is
\[
  \tau=\frac{\theta}{2\pi}+\frac{4\pi\ii}{g_{\rm YM}^2}
\]
in the present trace convention.
\end{definition}

Equivalently, after choosing an \(\mathcal N=1\) subalgebra, the same field
field content is an \(\mathcal N=1\) vector multiplet plus three adjoint chiral
multiplets \(\Phi_1,\Phi_2,\Phi_3\), with cubic superpotential
\[
  W_{\mathcal N=4}
  =
  \sqrt2\,\operatorname{tr}\bigl(\Phi_1[\Phi_2,\Phi_3]\bigr),
\]
after the superpotential coupling has been tied to \(g_{\rm YM}\) by the
enhanced supersymmetry.  This \(\mathcal N=1\) presentation is a bookkeeping
device for an \(SU(3)\times U(1)\) subgroup of the full \(SU(4)_R\) theory.

\begin{proposition}[One-loop and holomorphic beta-function cancellation]
\label{prop:n4-beta-one-loop-cancellation}
Let \(C_2(\mathfrak g)\) denote the adjoint quadratic Casimir in the chosen
trace convention.  The one-loop gauge beta-function coefficient of the
\(\mathcal N=4\) field content vanishes.  In the \(\mathcal N=1\) Wilsonian
coordinate, the one-loop holomorphic coefficient vanishes as well.
\end{proposition}

\begin{proof}
In an ordinary four-dimensional gauge-theory convention, one vector, four
adjoint Weyl fermions, and six adjoint real scalars contribute
\[
  b_0
  =
  C_2(\mathfrak g)
  \left(
  \frac{11}{3}
  -4\frac{2}{3}
  -6\frac{1}{6}
  \right)
  =
  0 .
\]
In the \(\mathcal N=1\) presentation, the holomorphic Wilsonian coefficient
is
\[
  b_0^{\rm hol}
  =
  3\,C_2(\mathfrak g)-\sum_{a=1}^3 T({\rm adj})
  =
  3\,C_2(\mathfrak g)-3\,C_2(\mathfrak g)
  =
  0 .
\]
The two calculations use different coupling coordinates, but both make the
same elementary field-content cancellation explicit.
\end{proof}

\begin{hypothesis}[Finite \(\mathcal N=4\) conformal-family input]
\label{hyp:n4-finite-conformal-family}
Fix the field content of Definition~\ref{def:n4-yang-mills-field-content}.  Assume
there is a local regulator and renormalization prescription with the following
properties.
\begin{enumerate}
\item Gauge Ward identities, locality, Poincare covariance, all sixteen
  supercharges, and \(SU(4)_R\) Ward identities are implemented without an
  anomaly.
\item The renormalized stress tensor and the \(SU(4)_R\) currents belong to a
  conserved \(\mathfrak{psu}(2,2|4)\) current multiplet, and the operator
  sourced by the complex coupling \(\tau\) is the corresponding
  dimension-four top component.
\item After the gauge-field normalization and possible background contact
  terms are fixed, no additional \(\mathcal N=4\)-preserving marginal source
  or current-moment-map obstruction is generated in the \(\tau\) direction.
\item In this source coordinate the beta function of \(\tau\) vanishes.
\end{enumerate}
\end{hypothesis}

This hypothesis is the precise status boundary.  The one-loop calculation in
Proposition~\ref{prop:n4-beta-one-loop-cancellation} is an elementary check on
the required cancellation; the all-order assertion is a renormalization
theorem with regulator-level input, independent of the spin-chain
construction.

\begin{definition}[Coupling chart and theta periodicity]
\label{def:n4-coupling-chart}
The local Yang--Mills coupling chart is the upper half-plane
\[
  \mathfrak H=\{\tau\in\mathbb C\mid \operatorname{Im}\tau>0\},
  \qquad
  \tau=\frac{\theta}{2\pi}+\frac{4\pi\ii}{g_{\rm YM}^2}.
\]
When the trace form and global gauge choice are chosen so that
\[
  k[A]=\frac{1}{32\pi^2}\int\operatorname{tr}(F_{\mu\nu}\widetilde F^{\mu\nu})
  \in\mathbb Z
\]
on closed spin four-manifolds in the allowed bundle sectors, the path-integral
phase \(\exp(\ii\theta k[A])\) is unchanged by
\[
  T:\tau\longmapsto\tau+1 .
\]
Thus \(q=\exp(2\pi\ii\tau)\) is the natural weak-coupling coordinate near the
cusp \(\operatorname{Im}\tau=\infty\).
\end{definition}

\begin{proposition}[Local \(\mathcal N=4\) conformal manifold]
\label{prop:n4-local-conformal-manifold}
Assume Hypothesis~\ref{hyp:n4-finite-conformal-family}.  Before imposing any
duality identifications, the \(\mathcal N=4\) Yang--Mills coupling chart
\(\mathfrak H\) is a one-complex-dimensional local conformal manifold.  The
tangent vector represented by \(\partial/\partial\tau\) is the exactly
marginal deformation obtained by integrating the protected dimension-four
top component of the stress-tensor multiplet.
\end{proposition}

\begin{proof}
The source chart has one complex coordinate, namely \(\tau\).  By
Hypothesis~\ref{hyp:n4-finite-conformal-family}, the beta map in this chart
is identically zero after the declared normalization and contact-term choices
are fixed, and no redundant source direction remains to be quotiented.  The
zero locus of the beta map is therefore all of \(\mathfrak H\).  Its complex
dimension is
\[
  1-0=1 .
\]
The inserted operator \(\partial S/\partial\tau\) is exactly marginal because
it has protected dimension \(4\) and its source has zero beta function in the
same chart.  This proves the local statement.  Global electric-magnetic
identifications between different values of \(\tau\) are additional data.
\end{proof}

\begin{hypothesis}[Duality-group input]
\label{hyp:n4-duality-group-input}
For a fixed global form of the gauge group and a fixed line-operator lattice,
let \(\Gamma\subset PSL(2,\mathbb Z)\) be the subgroup of genuine
four-dimensional \(\mathcal N=4\) QFT equivalences acting on the same
extended theory structure.  Its elements act on the coupling by
\[
  \tau\longmapsto\frac{a\tau+b}{c\tau+d},
  \qquad
  \begin{pmatrix}a&b\\ c&d\end{pmatrix}\in SL(2,\mathbb Z).
\]
For the standard local-operator sector of \(SU(N)\) Yang--Mills one often
uses the full \(PSL(2,\mathbb Z)\) generated by
\[
  T:\tau\mapsto\tau+1,\qquad S:\tau\mapsto-\frac1{\tau},
\]
but the precise global quotient depends on the line-operator and global-form
structure.  This hypothesis is an input about nonperturbative QFT equivalences,
with no string-theory construction entering the argument.
\end{hypothesis}

\begin{proposition}[Global quotient, conditional on duality]
\label{prop:n4-duality-quotient}
Assume Hypotheses~\ref{hyp:n4-finite-conformal-family} and
\ref{hyp:n4-duality-group-input}.  The local conformal family
\(\mathfrak H\) descends, for the chosen extended theory structure, to the
orbifold conformal parameter space
\[
  \mathfrak H/\Gamma .
\]
The generator \(T\) is the theta-periodicity of
Definition~\ref{def:n4-coupling-chart}; the generator \(S\), when it belongs
to \(\Gamma\), exchanges the weak-coupling cusp with a strong-coupling region.
\end{proposition}

\begin{proof}
The action of \(SL(2,\mathbb Z)\) on \(\mathfrak H\) preserves
\(\operatorname{Im}\tau>0\) because
\[
  \operatorname{Im}\frac{a\tau+b}{c\tau+d}
  =
  \frac{\operatorname{Im}\tau}{|c\tau+d|^2}.
\]
Matrices differing by an overall sign have the same fractional-linear action,
so the effective group is a subgroup of \(PSL(2,\mathbb Z)\).  If
\(\gamma\in\Gamma\) is a genuine QFT equivalence for the fixed extended structure,
then \(\tau\) and \(\gamma\cdot\tau\) label the same point of the global
conformal parameter space.  Quotienting \(\mathfrak H\) by these
identifications gives \(\mathfrak H/\Gamma\).  The statement is conditional
on the duality hypothesis; local exact marginality alone gives only
\(\mathfrak H\).
\end{proof}

The spin-chain spectral problem below uses the local SCFT family and the
planar single-trace dilatation operator, with the global duality quotient
kept as separate input.

\begin{proposition}[Conformal anomaly coefficients]
\label{prop:n4-central-charges}
For \(G=SU(N)\),
\[
  a=c=\frac{\dim\mathfrak{su}(N)}{4}
      =\frac{N^2-1}{4}.
\]
For \(G=U(N)\), including the free abelian multiplet,
\[
  a=c=\frac{N^2}{4}.
\]
\end{proposition}

\begin{proof}
The trace-anomaly coefficients of free conformal fields in four dimensions
are
\[
  a_{\rm scalar}=\frac{1}{360},\quad
  a_{\rm Weyl}=\frac{11}{720},\quad
  a_{\rm vector}=\frac{31}{180},
\]
and
\[
  c_{\rm scalar}=\frac{1}{120},\quad
  c_{\rm Weyl}=\frac{1}{40},\quad
  c_{\rm vector}=\frac{1}{10}.
\]
One \(\mathcal N=4\) vector multiplet per generator of \(\mathfrak g\)
contains six real scalars, four Weyl fermions, and one gauge vector.  Hence
\[
  a_{\rm multiplet}
  =
  6\frac{1}{360}
  +4\frac{11}{720}
  +\frac{31}{180}
  =
  \frac14,
\]
and
\[
  c_{\rm multiplet}
  =
  6\frac{1}{120}
  +4\frac{1}{40}
  +\frac{1}{10}
  =
  \frac14 .
\]
Multiplying by \(\dim\mathfrak g\) gives the displayed formulae.  The use of
the free-field coefficients is justified because \(a\) and \(c\) sit in the
same protected stress-tensor multiplet data along the exactly marginal
\(\mathcal N=4\) conformal manifold.
\end{proof}

\section{Protected Operators and Elementary OPE Data}
\label{sec:n4-protected-operators-ope-data}

Let \(C_{I_1\cdots I_k}\) be a symmetric traceless tensor of \(SO(6)_R\):
\[
  C_{I_{\sigma(1)}\cdots I_{\sigma(k)}}=C_{I_1\cdots I_k},
  \qquad
  \delta^{I_aI_b}C_{I_1\cdots I_k}=0 .
\]
The corresponding single-trace half-BPS representative is
\[
  \mathcal O_C^{(k)}
  =
  C_{I_1\cdots I_k}\,
  \operatorname{tr}\!\bigl(
  \Phi^{I_1}\cdots\Phi^{I_k}
  \bigr),
\]
with normal ordering and operator mixing understood in the renormalized
local-operator space.  It transforms in the \(SU(4)_R\) representation
\([0,k,0]\).

\begin{quotedtheorem}[Half-BPS protection]
\label{qthm:n4-half-bps-protection}
In unitary four-dimensional \(\mathcal N=4\) superconformal Yang--Mills
theory, a scalar superconformal primary in the half-BPS multiplet
\([0,k,0]\) has exact scaling dimension
\[
  \Delta=k.
\]
The two- and three-point functions of half-BPS primaries, after fixing their
normalization, are independent of the exactly marginal coupling.
\end{quotedtheorem}

This is a representation-theoretic and nonrenormalization statement.  It does
not say that every gauge-invariant word with classical dimension \(k\) is
protected.  The singlet
\[
  \mathcal K=\operatorname{tr}(\Phi^I\Phi^I)
\]
is the Konishi primary; it belongs to a long multiplet and acquires an
anomalous dimension.  The traceless tensor
\[
  \mathcal O^{IJ}
  =
  \operatorname{tr}(\Phi^I\Phi^J)
  -\frac16\delta^{IJ}\operatorname{tr}(\Phi^K\Phi^K)
\]
is the bottom component of the stress-tensor multiplet and has
\(\Delta=2\).

\begin{proposition}[Stress-tensor-multiplet normalization]
\label{prop:n4-stress-tensor-multiplet-normalization}
Choose an orthonormal basis \(T^a\) of \(\mathfrak g\) with
\(\operatorname{tr}(T^aT^b)=\delta^{ab}\), and write
\[
  \langle\Phi^{I,a}(x)\Phi^{J,b}(0)\rangle_{\rm free}
  =
  \frac{g_{\rm YM}^2\,\delta^{IJ}\delta^{ab}}{4\pi^2 x^2}.
\]
Define
\[
  B^{IJ}=g_{\rm YM}^{-2}:\operatorname{tr}(\Phi^I\Phi^J):,
  \qquad
  O^{IJ}=B^{IJ}-\frac16\delta^{IJ}B^{KK},
\]
and let
\[
  P^{IJ,KL}
  =
  \frac12(\delta^{IK}\delta^{JL}+\delta^{IL}\delta^{JK})
  -\frac16\delta^{IJ}\delta^{KL}
\]
be the projector onto symmetric traceless \(SO(6)\) tensors.  Then
\[
  \langle O^{IJ}(x)O^{KL}(0)\rangle_{\rm free}
  =
  \frac{\dim\mathfrak g}{8\pi^4 x^4}\,P^{IJ,KL}.
\]
Consequently
\[
  \mathbb O^{IJ}
  =
  \frac{2\sqrt2\,\pi^2}{\sqrt{\dim\mathfrak g}}\,O^{IJ}
\]
has the canonical two-point function
\[
  \langle\mathbb O^{IJ}(x)\mathbb O^{KL}(0)\rangle
  =
  \frac{P^{IJ,KL}}{x^4}.
\]
\end{proposition}

\begin{proof}
Wick contraction gives
\[
  \langle B^{IJ}(x)B^{KL}(0)\rangle_{\rm free}
  =
  \frac{\dim\mathfrak g}{16\pi^4x^4}
  \bigl(\delta^{IK}\delta^{JL}+\delta^{IL}\delta^{JK}\bigr).
\]
Subtracting the trace in each pair projects both index pairs with \(P\).  The
result is
\[
  \frac{\dim\mathfrak g}{16\pi^4x^4}
  \left(
  \delta^{IK}\delta^{JL}
  +\delta^{IL}\delta^{JK}
  -\frac13\delta^{IJ}\delta^{KL}
  \right)
  =
  \frac{\dim\mathfrak g}{8\pi^4x^4}P^{IJ,KL}.
\]
Multiplication by \((2\sqrt2\pi^2)^2/\dim\mathfrak g\) gives the normalized
two-point function.
\end{proof}

There is also a useful planar chiral-ring convention.  For \(G=SU(N)\), let
\[
  Z=\frac{1}{\sqrt2}(\Phi^5+\ii\Phi^6),
  \qquad
  \mathcal Z_J
  =
  \frac{(2\pi)^J}{g_{\rm YM}^J\sqrt{J\,N^J}}\,
  :\operatorname{tr}(Z^J):
\]
in the leading large-\(N\) single-trace normalization.

\begin{proposition}[Planar chiral-primary normalization]
\label{prop:planar-n4-chiral-primary-normalization}
At leading planar order and for fixed \(J,J_1,J_2\) as \(N\to\infty\),
\[
  \langle \mathcal Z_J(x)\,\overline{\mathcal Z}_J(0)\rangle_{\rm pl}
  =
  \frac{1}{x^{2J}},
\]
and, for \(J=J_1+J_2\),
\[
  \bigl\langle
  \mathcal Z_{J_1}(x_1)\mathcal Z_{J_2}(x_2)
  \overline{\mathcal Z}_{J}(x_3)
  \bigr\rangle_{\rm pl}
  =
  \frac{\sqrt{J_1J_2J}}{N}\,
  \frac{1}{x_{13}^{2J_1}x_{23}^{2J_2}} .
\]
Equivalently, in the protected chiral OPE,
\[
  \mathcal Z_{J_1}(x)\mathcal Z_{J_2}(0)
  =
  \frac{\sqrt{J_1J_2(J_1+J_2)}}{N}\,
  \mathcal Z_{J_1+J_2}(0)
  +\text{descendants and multi-traces}
\]
at leading planar order.
\end{proposition}

\begin{proof}
The elementary propagator in the trace convention used above is
\[
  \langle Z^a(x)\overline Z^b(0)\rangle_{\rm free}
  =
  \frac{g_{\rm YM}^2}{4\pi^2x^2}\delta^{ab}.
\]
For a single trace of length \(J\), planar Wick contractions pair each \(Z\)
with the corresponding \(\overline Z\) in a cyclic order.  There are \(J\)
choices of the relative cyclic origin, and the double-line color graph has
\(J\) leading color loops.  Hence
\[
  \bigl\langle :\operatorname{tr}(Z^J):(x)
  :\operatorname{tr}(\overline Z^J):(0)\bigr\rangle_{\rm pl}
  =
  JN^J
  \left(\frac{g_{\rm YM}^2}{4\pi^2x^2}\right)^J .
\]
Multiplying by the two normalization factors in \(\mathcal Z_J\) gives the
unit two-point function.

For the extremal three-point function put \(J=J_1+J_2\).  At leading planar
order all fields in the first two holomorphic traces contract with the single
anti-holomorphic trace.  A planar embedding is fixed by choosing the cyclic
origin of the anti-holomorphic trace, the cyclic origin of the
\(J_1\)-trace, and the cyclic origin of the \(J_2\)-trace.  This gives
\(JJ_1J_2\) planar embeddings.  The connected double-line graph has
\(J-1\) leading color loops, because three trace boundaries have been joined
into one planar pair-of-pants graph.  Therefore
\[
\begin{aligned}
  &\bigl\langle
  :\operatorname{tr}(Z^{J_1}):(x_1)
  :\operatorname{tr}(Z^{J_2}):(x_2)
  :\operatorname{tr}(\overline Z^J):(x_3)
  \bigr\rangle_{\rm pl}
\\
  &\hspace{4em}
  =
  JJ_1J_2\,N^{J-1}
  \left(\frac{g_{\rm YM}^2}{4\pi^2}\right)^J
  \frac{1}{x_{13}^{2J_1}x_{23}^{2J_2}} .
\end{aligned}
\]
The product of the three normalization constants contributes
\[
  \frac{(2\pi)^{2J}}
       {g_{\rm YM}^{2J}\sqrt{J_1N^{J_1}}\sqrt{J_2N^{J_2}}\sqrt{JN^J}},
\]
which reduces the coefficient to \(\sqrt{J_1J_2J}/N\).  The displayed OPE
coefficient follows by taking \(x_1\to x_2\) in this protected extremal
three-point function with the two-point normalization already fixed.
\end{proof}

Finite-\(N\) trace relations and multi-trace mixing are not suppressed in
this formula; they are part of the exact protected-operator problem rather
than the single-trace spin-chain problem.

\section{Planar Limit and Operator Hilbert Space}

Let the gauge group be \(SU(N)\) or \(U(N)\), with gauge coupling defined by
the kinetic term
\[
  \frac{1}{4g_{\rm YM}^2}\int \dd^4x\,
  \operatorname{tr}(F_{\mu\nu}F^{\mu\nu})
\]
in the trace normalization used throughout the gauge-theory volumes.  The
't Hooft coupling is
\[
  \lambda=g_{\rm YM}^2N,\qquad
  g=\frac{\sqrt\lambda}{4\pi}.
\]
The planar limit means
\[
  N\to\infty,\qquad \lambda\ \text{fixed},
\]
with local single-trace operators normalized so that their planar two-point
functions remain finite.

Let \(\mathfrak V\) denote the one-letter vector space spanned by the
elementary adjoint fields of \(\mathcal N=4\) SYM and their covariant
derivatives, modulo the equations of motion and gauge-covariant descendant
relations appropriate to local-operator renormalization.  For fixed length
\(L\), a single-trace word is
\[
  \mathcal O_W(x)
  =
  \operatorname{tr}\!\bigl(W_1(x)W_2(x)\cdots W_L(x)\bigr),
  \qquad W_i\in\mathfrak V .
\]
Cyclicity of the trace identifies
\[
  W_1\otimes W_2\otimes\cdots\otimes W_L
  \sim
  W_2\otimes\cdots\otimes W_L\otimes W_1 .
\]
Thus the fixed-length single-trace space is the cyclic quotient
\[
  \mathcal H_L^{\rm cyc}
  =
  \mathfrak V^{\otimes L}/\langle\tau-1\rangle ,
\]
where \(\tau\) shifts all letters by one site.
For planar perturbation theory this quotient is equipped with the large-\(N\)
inner product obtained from the leading two-point function.  If
\(W=W_1\otimes\cdots\otimes W_L\) and \(W'=W'_1\otimes\cdots\otimes W'_L\)
are words of the same classical dimension, define
\[
  \langle W,W'\rangle_{\rm pl}
  =
  \frac{1}{L\,N^L}
  \sum_{r=0}^{L-1}
  \prod_{j=1}^{L}
  \langle W_j,W'_{j+r}\rangle_{\rm letter},
  \qquad j+r\ {\rm read\ modulo}\ L .
\]
The factor \(N^{-L}\) removes the leading color loop, while \(L^{-1}\)
averages over the trace origin.  Operators differing by equations of motion,
gauge-covariant total derivatives, or \(Q\)-exact representatives in a
chosen protected cohomological sector are first quotient out in the
renormalized local-operator space; the spin-chain word is a representative of
that quotient, not a new fundamental Hilbert space.

\begin{definition}[Planar single-trace spectral problem]
\label{def:planar-single-trace-spectral-problem}
Fix a renormalization scheme preserving the manifest \(SU(4)_R\) and
Poincare symmetries of \(\mathcal N=4\) SYM and fix a closed sector
\(\mathcal S\subset \mathfrak V\) under planar mixing to a specified loop
order.  The planar single-trace spectral problem of length \(L\) is the cyclic
quotient \(\mathcal S^{\otimes L}/\langle\tau-1\rangle\), its planar
two-point inner product above, and the restriction of the anomalous-dimension
matrix \(\Gamma\) to the quotient.  A spin-chain Hamiltonian is a coordinate
description of this problem after a basis of letters has been chosen.
\end{definition}

The dilatation operator \(D\) is defined by the logarithmic scale dependence
of renormalized local operators.  In a basis \(\mathcal O_A\),
\[
  \mu\frac{\dd}{\dd\mu}\mathcal O_A
  =
  -\Gamma_A{}^B\mathcal O_B,\qquad
  D_A{}^B=(D_0)_A{}^B+\Gamma_A{}^B .
\]
At a conformal fixed point, diagonalizing \(D\) is equivalent to finding local
operators of definite scaling dimension.  Planarity implies that, at fixed
loop order, \(D\) acts locally along the cyclic word: at \(\ell\) loops the
range is at most \(\ell+1\) adjacent sites in the scalar subsectors described
below.

\section{The SU(2) Scalar Sector}

Choose two complex scalar fields
\[
  Z=\frac{1}{\sqrt2}(\phi^5+\ii\phi^6),
  \qquad
  X=\frac{1}{\sqrt2}(\phi^1+\ii\phi^2).
\]
The \(SU(2)\) scalar sector is spanned by
\[
  \operatorname{tr}(W_1\cdots W_L),\qquad W_i\in\{Z,X\}.
\]
Identify \(Z\) with a down spin and \(X\) with an up spin:
\[
  Z\leftrightarrow\ket{\downarrow},\qquad
  X\leftrightarrow\ket{\uparrow}.
\]
Let \(P_{j,j+1}\) be the operator that interchanges the two letters at sites
\(j\) and \(j+1\), with \(L+1\equiv1\).

The local mixing calculation may be reduced to a two-site problem.  In the
scalar potential, the term relevant for holomorphic \(X,Z\) words is
\[
  V_{\rm sc}
  =
  -\frac{1}{2g_{\rm YM}^2}
  \operatorname{tr}\bigl([\Phi_I,\Phi_J][\Phi_I,\Phi_J]\bigr)
\]
with the complex fields embedded in the six real scalars \(\Phi_I\).  In
double-line notation, a planar contraction of a quartic vertex with two
adjacent letters has one color loop fewer than the free two-point function
and hence contributes \(g_{\rm YM}^2N=\lambda\).  Non-adjacent contractions
are non-planar in a single trace and are suppressed by \(N^{-2}\).

\begin{proposition}[Two-site scalar mixing tensor]
\label{prop:planar-n4-two-site-mixing-tensor}
On adjacent holomorphic \(SU(2)\) letters, the logarithmic one-loop mixing
tensor has the form
\[
  \Gamma_{j,j+1}^{(1)}
  =
  \frac{\lambda}{8\pi^2}\,(1-P_{j,j+1})
\]
after adding the scalar self-energy and gauge-exchange contributions required
by \(\mathcal N=4\) supersymmetry.  The statement is perturbative in the
chosen dimensional-regularization/minimal-subtraction convention.
\end{proposition}

\begin{proof}
The quartic vertex gives a local tensor generated by the two invariant maps
on two \(SU(2)\) letters: the identity \(I\) and the permutation \(P\).  The
planar color contraction of
\(\operatorname{tr}(T^{a_1}T^{a_2}\cdots T^{a_L})\) with a vertex joining
sites \(j,j+1\) produces \(N\operatorname{tr}(\cdots T^{b_j}T^{b_{j+1}}\cdots)\)
at leading order, so the coefficient is proportional to \(\lambda\) and no
operator with separated sites appears.  The scalar bubble integral in
\(d=4-2\varepsilon\) dimensions has logarithmic part
\[
  \frac{1}{(4\pi)^2}\frac{1}{\varepsilon}
  \quad\Longleftrightarrow\quad
  \frac{1}{8\pi^2}\log\mu
\]
in the anomalous-dimension convention used for two-point functions.  The
exchange part is \(-P\).  The identity part receives the non-exchange scalar,
gauge, and self-energy pieces.  The protected operator
\(\operatorname{tr}(Z^L)\) is an eigenvector with zero anomalous dimension,
so the total two-site tensor annihilates every symmetric \(ZZ\) pair.  This
fixes the identity coefficient to \(+1\) in the same normalization, giving
\((\lambda/8\pi^2)(I-P)\).
\end{proof}

\section{The SO(6) Scalar Hamiltonian and the SU(2) Reduction}

The \(SU(2)\) calculation is a subsector of the full scalar one-loop
Hamiltonian.  Let the six real scalars be denoted \(\Phi_I\),
\(I=1,\ldots,6\), and write a two-site scalar state as
\(\Phi_I\otimes\Phi_J\).  There are three invariant maps on
\(\mathbb R^6\otimes\mathbb R^6\):
\[
\begin{aligned}
  I(\Phi_I\otimes\Phi_J)&=\Phi_I\otimes\Phi_J,\\
  P(\Phi_I\otimes\Phi_J)&=\Phi_J\otimes\Phi_I,\\
  K(\Phi_I\otimes\Phi_J)&=\delta_{IJ}\sum_{K=1}^6\Phi_K\otimes\Phi_K .
\end{aligned}
\]
Here \(P\) permutes the two \(SO(6)\) vector indices and \(K\) is the trace
operator.  In the normalization of
Theorem~\ref{thm:planar-n4-one-loop-su2}, the scalar one-loop nearest-neighbor
density is
\begin{equation}
  \Gamma^{(1)}_{j,j+1}
  =
  \frac{\lambda}{16\pi^2}\bigl(K_{j,j+1}+2I_{j,j+1}-2P_{j,j+1}\bigr).
  \label{eq:planar-n4-so6-one-loop-density}
\end{equation}
The \(K\)-term is the place where the full \(SO(6)\) chain differs from the
holomorphic \(SU(2)\) chain.

\begin{proposition}[Holomorphic reduction of the \(SO(6)\) density]
\label{prop:planar-n4-so6-to-su2}
On the holomorphic sector generated by \(Z\) and \(X\), the trace operator
\(K\) in \eqref{eq:planar-n4-so6-one-loop-density} vanishes, and the density
reduces to
\[
  \frac{\lambda}{8\pi^2}(I-P).
\]
\end{proposition}

\begin{proof}
The operator \(K\) contracts the \(SO(6)\) indices of the two adjacent scalar
letters.  In the complex basis \(Z=(\Phi_5+\ii\Phi_6)/\sqrt2\) and
\(X=(\Phi_1+\ii\Phi_2)/\sqrt2\), the holomorphic contractions are null:
\[
  Z\cdot Z=X\cdot X=Z\cdot X=0 .
\]
Thus \(K\) annihilates \(Z\otimes Z\), \(Z\otimes X\), \(X\otimes Z\), and
\(X\otimes X\).  Substituting \(K=0\) in
\eqref{eq:planar-n4-so6-one-loop-density} gives
\[
  \frac{\lambda}{16\pi^2}(2I-2P)
  =
  \frac{\lambda}{8\pi^2}(I-P).
\]
\end{proof}

This reduction is a useful convention check.  If the overall coefficient or
the sign of \(P\) is changed, the \(SO(6)\) formula no longer restricts to the
Konishi normalization computed below.

\begin{theorem}[One-loop planar \(SU(2)\) Hamiltonian]
\label{thm:planar-n4-one-loop-su2}
In the \(SU(2)\) scalar sector, with the coupling convention above, the planar
one-loop dilatation operator is
\[
  D
  =
  L+\frac{\lambda}{8\pi^2}H_{XXX}
  +O(\lambda^2),
  \qquad
  H_{XXX}=\sum_{j=1}^{L}(1-P_{j,j+1})
\]
on the cyclic spin-chain quotient.
\end{theorem}

\begin{proof}
The only planar one-loop terms that move \(X\) and \(Z\) letters inside the
holomorphic \(SU(2)\) sector come from the scalar quartic interaction and the
wavefunction and non-exchange diagrams required by supersymmetry.  The
exchange part is local on adjacent sites and is proportional to
\(-P_{j,j+1}\).  The non-exchange part is proportional to the identity on the
same pair.  Hence the most general nearest-neighbor planar one-loop operator
compatible with the two-letter \(SU(2)\) symmetry and translation around the
trace is
\[
  c\sum_{j=1}^L(1-P_{j,j+1})+c'\sum_{j=1}^L(1+P_{j,j+1}).
\]
The half-BPS operators \(\operatorname{tr}(Z^L)\) and their fully symmetrized
scalar descendants have protected scaling dimension.  On a symmetric pair,
\(P_{j,j+1}=1\), so protection forces \(c'=0\).  The coefficient \(c\) is
fixed by the logarithmic part of the single adjacent exchange integral.  In
the present normalization the divergent scalar integral is
\[
  \frac{\lambda}{8\pi^2}\log(\mu |x|)+\text{finite},
\]
and its color contraction is planar and nearest-neighbor.  Substituting this
coefficient gives the displayed \(H_{XXX}\).  The cyclic quotient implements
the trace relation and imposes zero total spin-chain momentum on physical
single-trace states.
\end{proof}

The proof uses supersymmetric protection only to fix the identity term after
the exchange coefficient has been computed.  This is a useful place where the
field-theory origin of the spin chain matters: the spin-chain Hamiltonian is
not introduced independently of the renormalized local-operator problem.

\section{Coordinate Bethe Ansatz}

We now diagonalize the dimensionless Hamiltonian
\[
  H_{XXX}=\sum_{j}(1-P_{j,j+1})
\]
before reinstating the overall factor \(\lambda/(8\pi^2)\).  On an infinite
chain let \(\ket{n}\) denote the state with a single \(X\) at site
\(n\in\mathbb Z\) and \(Z\)'s elsewhere.  A one-magnon wavefunction is
\[
  \Psi_p(n)=\ee^{\ii pn}.
\]
The Hamiltonian \(H_{XXX}\) gives
\[
  (H_{XXX}\Psi_p)(n)
  =
  2\Psi_p(n)-\Psi_p(n+1)-\Psi_p(n-1),
\]
hence
\[
  E_1(p)=4\sin^2\frac p2 .
\]
The corresponding one-loop anomalous dimension contribution is
\[
  \gamma_1(p)=\frac{\lambda}{8\pi^2}E_1(p).
\]
On a cyclic single trace a single magnon with nonzero momentum is not a
physical operator, because trace cyclicity requires total momentum zero.  The
one-magnon calculation is nevertheless the local dispersion input for
multi-magnon states.

\begin{proposition}[Two-magnon matching for the one-loop chain]
\label{prop:planar-n4-two-magnon-matching}
Let \(z_a=\ee^{\ii p_a}\), \(a=1,2\), and let
\[
  u_a=\frac12\cot\frac{p_a}{2},
  \qquad
  z_a=\frac{u_a+\ii/2}{u_a-\ii/2}.
\]
In the two-magnon chamber \(n_1<n_2\), set
\[
  \Psi(n_1,n_2)
  =
  A_{12}\ee^{\ii(p_1n_1+p_2n_2)}
  +
  A_{21}\ee^{\ii(p_2n_1+p_1n_2)} .
\]
Then \(\Psi\) is an eigenfunction of \(H_{XXX}\) on the infinite chain with
energy \(E_1(p_1)+E_1(p_2)\) precisely when
\[
  \frac{A_{21}}{A_{12}}
  =
  \mathsf a(p_1,p_2)
  =
  -\frac{1-2\ee^{\ii p_2}+\ee^{\ii(p_1+p_2)}}
          {1-2\ee^{\ii p_1}+\ee^{\ii(p_1+p_2)}}
  =
  \frac{u_1-u_2-\ii}{u_1-u_2+\ii}.
\]
\end{proposition}

\begin{proof}
When the magnons are separated, \(n_2>n_1+1\), the Hamiltonian is the sum of
two one-particle finite-difference operators:
\[
\begin{aligned}
  (H_{XXX}\Psi)(n_1,n_2)
  &=
  4\Psi(n_1,n_2)-\Psi(n_1-1,n_2)-\Psi(n_1+1,n_2)
\\
  &\quad
  -\Psi(n_1,n_2-1)-\Psi(n_1,n_2+1).
\end{aligned}
\]
Hence the separated equation fixes the energy to
\(E_1(p_1)+E_1(p_2)\).  The only remaining condition is the contact equation
at \(n_2=n_1+1\).  There the adjacent \(XX\) pair is symmetric, so the
middle link contributes zero and only the two exterior links move a magnon:
\[
  (H_{XXX}\Psi)(n,n+1)
  =
  2\Psi(n,n+1)-\Psi(n-1,n+1)-\Psi(n,n+2).
\]
Substituting the plane-wave ansatz and the separated energy gives
\[
  0
  =
  -A_{21}z_1z_2+2A_{21}z_1-A_{21}
  -A_{12}z_1z_2+2A_{12}z_2-A_{12}.
\]
Solving this linear equation for \(A_{21}/A_{12}\) gives the rational
expression in \(z_1,z_2\).  Rewriting \(z_a=(u_a+\ii/2)/(u_a-\ii/2)\) gives
the displayed rapidity form.
\end{proof}

The coefficient \(\mathsf a(p_1,p_2)\) is the exchange coefficient for the
ordered chamber convention just chosen.  The Bethe--Yang phase in the closed
chain is its inverse:
\[
  S_{\rm BY}(u_j,u_k)
  =
  \mathsf a(p_j,p_k)^{-1}
  =
  \frac{u_j-u_k+\ii}{u_j-u_k-\ii}.
\]
This inverse is a convention-sensitive point.  For two magnons, trace
periodicity says
\[
  \Psi(n_1,n_2)=\Psi(n_2,n_1+L),\qquad n_1<n_2.
\]
Comparing the coefficient of
\(\ee^{\ii(p_1n_1+p_2n_2)}\) gives
\[
  A_{12}=A_{21}\ee^{\ii p_1L},
  \qquad
  \ee^{\ii p_1L}=S_{\rm BY}(u_1,u_2).
\]
The second equation is obtained by moving the other magnon around the trace.
For \(M\) magnons the one-loop \(XXX_{1/2}\) chain has factorized scattering;
equivalently, the transfer matrices constructed in
Chapter~\ref{ch:integrable-algebraic-bethe-ansatz-transfer-matrices}
commute and the coordinate wavefunction is consistent in every ordering
chamber.  The finite periodic-chain equations are therefore
\begin{equation}
  \left(\frac{u_j+\ii/2}{u_j-\ii/2}\right)^L
  =
  \prod_{\substack{k=1\\k\neq j}}^M
  \frac{u_j-u_k+\ii}{u_j-u_k-\ii},
  \qquad
  \prod_{j=1}^M\frac{u_j+\ii/2}{u_j-\ii/2}=1 .
  \label{eq:planar-n4-one-loop-su2-bethe}
\end{equation}
The second equation is the cyclicity condition.  The energy is
\[
  E=\sum_{j=1}^M\frac{1}{u_j^2+1/4}.
\]

At one loop these Bethe equations are exact for the finite \(XXX_{1/2}\)
Hamiltonian.  The word ``asymptotic'' in later all-loop chapters has a
different meaning: the long-range planar dilatation operator ceases to be
finite-neighbor once the loop order grows, and finite trace length then
requires wrapping corrections.

\section{Konishi at One Loop}

The shortest non-BPS state in this sector is the length-four two-magnon
descendant represented by \(\operatorname{tr}([X,Z]^2)\).  Set
\[
  L=4,\qquad M=2,\qquad
  u_1=\frac{1}{2\sqrt3},\qquad u_2=-\frac{1}{2\sqrt3}.
\]
Then
\[
  \frac{u_1+\ii/2}{u_1-\ii/2}=\ee^{2\pi\ii/3},
  \qquad
  \frac{u_2+\ii/2}{u_2-\ii/2}=\ee^{-2\pi\ii/3},
\]
so the cyclicity product is \(1\).  The first Bethe equation reads
\[
  \ee^{8\pi\ii/3}
  =
  \frac{u_1-u_2+\ii}{u_1-u_2-\ii}
  =
  \ee^{2\pi\ii/3},
\]
and the second is its inverse.  The spin-chain energy is
\[
  E=\frac{1}{1/12+1/4}+\frac{1}{1/12+1/4}=6 .
\]
Therefore the one-loop anomalous dimension is
\[
  \gamma_{\rm K}^{(1)}
  =
  \frac{\lambda}{8\pi^2}\,6
  =
  \frac{3\lambda}{4\pi^2}.
\]

\section{Closed Sectors and Long-Range Data}

The \(SU(2)\) sector is the smallest laboratory, but the planar spectral
problem uses a hierarchy of closed sectors.  The \(SL(2)\) sector is spanned
by
\[
  \operatorname{tr}(D_+^{s_1}Z\,D_+^{s_2}Z\cdots D_+^{s_L}Z),
  \qquad s_j\in\mathbb Z_{\geq0},
\]
where \(D_+\) is a fixed lightlike covariant derivative.  Its one-loop chain
has noncompact spin at each site.  The \(SU(2|3)\) sector contains three
complex scalars and two fermions and is the smallest sector in which the
dynamic length-changing character of the higher-loop planar chain is visible.
The full one-letter module is the oscillator representation of
\(\mathfrak{psu}(2,2|4)\), and a local operator is a cyclic tensor product of
that module subject to gauge-invariant trace cyclicity.

At \(\ell\) loops in the asymptotic regime, the planar dilatation operator has
the range-\((\ell+1)\) asymptotic expansion
\[
  D
  =
  D_0+\sum_{\ell\geq1}g^{2\ell}
  \sum_{j=1}^{L}
  H^{(\ell)}_{j,\ldots,j+\ell},
  \qquad g=\frac{\sqrt\lambda}{4\pi}.
\]
For example, the \(SU(2)\) long-range Hamiltonian begins
\[
  H^{(1)}=2\{\,\}-2\{1\},
\]
in the permutation-structure notation where \(\{\,\}\) is the identity sum and
\(\{1\}=\sum_jP_{j,j+1}\).  Higher orders contain \(\{1,2\}\),
\(\{2,1\}\), and longer permutation words.  Their coefficients are not fixed
by \(SU(2)\) symmetry alone; they are constrained by Feynman graphs,
BMN-scaling tests, integrability, and supersymmetry.  The all-loop Bethe
ansatz in the next chapter should therefore be read as an asymptotic
integrability reconstruction of this long-range operator, not as an
independent definition of the four-dimensional theory.

\section{Beyond One Loop}

At higher loop order the planar dilatation operator becomes a long-range spin
chain.  In the \(SU(2)\) sector its range grows with loop order, and for an
operator of finite length \(L\) interactions can wrap around the trace once
the perturbative range reaches \(L\).  Thus there are two different spectral
problems:
\[
  \text{asymptotic long chain}
  \qquad\text{and}\qquad
  \text{finite trace with wrapping}.
\]
The next chapter describes the asymptotic all-loop Bethe ansatz.  The
following chapter explains how mirror TBA and the Y-system incorporate
finite-length effects.
