\chapter{Nonrenormalization and Holomorphy}
\label{ch:susy-nonrenormalization-holomorphy}

\ConventionsLorentzian

Nonrenormalization arguments are statements about Wilsonian coordinates in a
scheme with declared supersymmetry, gauge symmetry, locality, and anomaly
data.  This chapter formulates the perturbative superpotential argument, the
holomorphic gauge-coupling coordinate, and the passage to the NSVZ
beta-function coordinate in the language of
Chapter~\ref{ch:susy-wilsonian-schemes-exact-dynamics}.

\section{Chiral Local Coordinates}

Let \(\Phi^i\) be chiral superfields and \(t^A\) chiral coupling coordinates.
A Wilsonian chiral coordinate is a local functional of the form
\[
  I_F[\Phi;t]
  =
  \int d^4x\,d^2\theta\,
  \mathcal W(\Phi;t)+\text{c.c.}
\]
defined modulo \(\bar D_{\dot\alpha}\)-exact full-superspace terms and
equations of motion appropriate to the chosen Wilsonian action.  A full
superspace coordinate is
\[
  I_D[\Phi,\bar\Phi;t,\bar t]
  =
  \int d^4x\,d^4\theta\,
  \mathcal K(\Phi,\bar\Phi;t,\bar t).
\]
The distinction is cohomological: an \(F\)-term is represented by a chiral
integrand and the chiral measure, while a \(D\)-term is represented by the
full superspace measure.

\section{Perturbative Superpotential Argument}

\begin{proposition}[Perturbative Wilsonian superpotential argument]
\label{prop:perturbative-wilsonian-superpotential}
Assume a massive or Wilsonian infrared-regulated \(\mathcal N=1\) theory in a
holomorphic Wilsonian scheme.  Perturbative loop corrections do not generate
new superpotential terms beyond those already present as holomorphic local
coordinates, except through tree-level elimination of massive chiral fields
and through explicitly declared anomaly-induced coordinate changes.
\end{proposition}

\begin{proof}
In supergraph perturbation theory each internal propagator carries spinor
derivatives and each interaction vertex is integrated over either \(d^4\theta\)
or over a chiral measure that may be represented inside full superspace by
spinor derivatives acting on delta functions.  For a loop correction to
produce an \(F\)-term, all but one pair of Grassmann integrations would have
to be saturated while leaving only \(d^2\theta\) for the external local
functional.  The spinor-derivative algebra in a loop supplies the full
superspace measure unless an infrared singularity converts a nonlocal
full-superspace expression into a local chiral expression.  A Wilsonian
effective action at scale \(\Lambda\) excludes such singularities by
integrating out only modes above \(\Lambda\).  Therefore loop corrections are
full-superspace local coordinates.  Tree-level elimination of massive chiral
fields is different: it substitutes the solution of a chiral equation of
motion into the existing superpotential and hence stays within the chiral
coordinate class.  Anomaly-induced coordinate changes are precisely the
declared exceptions in the scheme data.
\end{proof}

\begin{proposition}[Tree-level elimination of a massive chiral block]
\label{prop:tree-level-holomorphic-elimination}
Let the chiral coordinates split as \((X^a,\Phi^i)\), where the \(X^a\) are
massive in a Wilsonian patch.  Suppose the holomorphic equations
\[
  \partial_{X^a}W(X,\Phi;t)=0
\]
have, on this patch, a unique holomorphic solution
\[
  X^a=X_*^a(\Phi;t),
\]
and suppose the heavy-block Hessian
\[
  H_{ab}=\partial_{X^a}\partial_{X^b}W(X_*,\Phi;t)
\]
is invertible there.  Then the tree-level eliminated superpotential
\[
  W_{\mathrm{eff}}(\Phi;t)=W(X_*(\Phi;t),\Phi;t)
\]
is holomorphic in the remaining chiral coordinates and couplings, and its
first derivatives are
\[
  \partial_{\Phi^i}W_{\mathrm{eff}}
  =
  \left.
  \partial_{\Phi^i}W
  \right|_{X=X_*}.
  \label{eq:tree-level-holomorphic-elimination-derivative}
\]
In the quadratic special case
\[
  W(X,\Phi;t)
  =
  W_0(\Phi;t)
  +J_a(\Phi;t)X^a
  +\frac12 M_{ab}(\Phi;t)X^aX^b,
  \qquad \det M\neq0,
\]
one has
\[
  X_*^a=-(M^{-1})^{ab}J_b,
  \qquad
  W_{\mathrm{eff}}
  =
  W_0-\frac12J_a(M^{-1})^{ab}J_b .
  \label{eq:quadratic-chiral-elimination}
\]
\end{proposition}

\begin{proof}
The hypotheses are precisely those of the holomorphic implicit-function
theorem applied to the map
\[
  F_a(X,\Phi;t)=\partial_{X^a}W(X,\Phi;t).
\]
Invertibility of \(H_{ab}=\partial F_a/\partial X^b\) at the solution gives
a unique holomorphic solution \(X_*(\Phi;t)\) after possibly shrinking the
patch.  Composing the holomorphic functions \(W\) and \(X_*\) gives a
holomorphic \(W_{\mathrm{eff}}\).  The chain rule gives
\[
  \partial_{\Phi^i}W_{\mathrm{eff}}
  =
  \left.
  \partial_{\Phi^i}W
  \right|_{X=X_*}
  +
  \left.
  \partial_{X^a}W
  \right|_{X=X_*}
  \partial_{\Phi^i}X_*^a .
\]
The second term vanishes by the heavy-field equation of motion, proving
\eqref{eq:tree-level-holomorphic-elimination-derivative}.

For the quadratic representative, the heavy-field equation is
\[
  J_a+M_{ab}X^b=0,
\]
so \(X_*^a=-(M^{-1})^{ab}J_b\).  Substitution gives
\[
  W_{\mathrm{eff}}
  =
  W_0
  -J_a(M^{-1})^{ab}J_b
  +\frac12M_{ab}(M^{-1})^{ac}J_c(M^{-1})^{bd}J_d
  =
  W_0-\frac12J_a(M^{-1})^{ab}J_b ,
\]
where symmetry of \(M_{ab}=\partial_a\partial_bW\) is used in the last
step.  If \(\det M\) vanishes, the patch contains an additional massless
chiral coordinate and the elimination is no longer a Wilsonian tree-level
operation in this sense.
\end{proof}

\section{Spurion Symmetries}

Treat a coupling \(t^A\) as a background chiral superfield.  If an exact
symmetry acts by
\[
  \Phi^i\mapsto e^{\ii q^i\alpha}\Phi^i,\qquad
  t^A\mapsto e^{\ii q^A\alpha}t^A,
\]
then a Wilsonian \(F\)-term must be a holomorphic section with the same total
charge as the chiral measure.  This turns dimensional analysis and
symmetries into a finite problem: list holomorphic monomials in fields and
couplings with the required charges and locality degree.  The conclusion is
only as strong as the declared symmetry and anomaly ledger.

\section{Holomorphic Gauge Coupling}

For a gauge theory with chiral matter, the holomorphic gauge coordinate is
\[
  \frac{1}{16\pi\ii}\int d^4x\,d^2\theta\,
  \tau(\Lambda)\operatorname{tr}(W^\alpha W_\alpha)+\text{c.c.}
\]
The trace convention is the one fixed in
Chapter~\ref{ch:susy-gauge-theory}.  Thus
\[
  b_0=3C_2(G)-\sum_iT(R_i)
\]
for chiral matter representations \(R_i\).  The indices are defined by
\[
  \operatorname{Tr}_{R_i}(XY)=T(R_i)\operatorname{tr}(XY),
  \qquad
  \operatorname{Tr}_{\rm ad}(XY)=C_2(G)\operatorname{tr}(XY).
\]

\begin{proposition}[Wilsonian one-loop exactness of the holomorphic gauge coordinate]
\label{prop:holomorphic-gauge-coupling-one-loop-exact}
Work in a weak-coupling Wilsonian patch in which:
\begin{enumerate}[label=(\roman*)]
\item the gauge kinetic coefficient is a chiral local coordinate
\(\tau_{\rm eff}(\mu)\), holomorphic in the microscopic chiral couplings;
\item the theta-angle periodicity is represented by
\(\tau\mapsto\tau+1\), so scale derivatives of the gauge kinetic coordinate
are single-valued holomorphic functions of
\[
  q_h:=\exp(2\pi\ii\tau);
\]
\item the perturbative sector is the \(q_h^0\) coefficient in the expansion
near \(q_h=0\); positive powers of \(q_h\) are instanton sectors and negative
powers are excluded from the weak-coupling Wilsonian patch;
\item the superpotential spurion ledger contains no additional anomalous
chiral coordinate with the transformation law of the gauge kinetic term.
\end{enumerate}
Then the perturbative Wilsonian running of the holomorphic coordinate is
\[
  \tau(\Lambda)
  =
  \tau(\Lambda_0)
  +\frac{b_0}{2\pi\ii}\log\frac{\Lambda}{\Lambda_0},
  \label{eq:holomorphic-gauge-coordinate-one-loop-running}
\]
up to a finite \(\Lambda\)-independent holomorphic redefinition of the
reference coordinate.  Equivalently,
\[
  \mu\frac{d\tau_{\rm eff}^{\rm pert}}{d\mu}
  =
  \frac{b_0}{2\pi\ii}.
\]
\end{proposition}

\begin{proof}
The statement is about a Wilsonian chiral coordinate, not about the canonical
two-point-function coupling.  By locality and chirality, the scale derivative
\(\mu\,d\tau_{\rm eff}/d\mu\) is itself the coefficient of the local chiral
operator \(\operatorname{tr}(W^\alpha W_\alpha)\).  It is dimensionless,
holomorphic in the chiral couplings, and invariant under
\(\tau\mapsto\tau+1\).  On the weak-coupling patch it therefore has an
expansion
\[
  \mu\frac{d\tau_{\rm eff}}{d\mu}
  =
  B_0(t)+\sum_{n\geq1}B_n(t)q_h^n ,
  \label{eq:holomorphic-gauge-beta-q-expansion}
\]
after excluding negative powers by the weak-coupling regularity hypothesis.
The \(n\geq1\) terms are nonperturbative instanton sectors.  The perturbative
projection is the constant term \(B_0(t)\).

The coefficient \(B_0(t)\) is constrained by the spurion symmetry ledger.
The gauge kinetic coordinate shifts under the anomalous chiral rotations of
charged fields; the local anomaly calculation fixes this shift at one loop
with coefficient \(b_0\).  A perturbative contribution depending
nontrivially on other chiral couplings would have to transform in the same
affine way as \(\tau\).  Hypothesis (iv) excludes such an additional
coordinate.  Hence \(B_0\) is the one-loop anomaly coefficient,
\[
  B_0=\frac{b_0}{2\pi\ii}.
\]
Integrating in \(\log\Lambda\) gives
\eqref{eq:holomorphic-gauge-coordinate-one-loop-running}.  A finite
holomorphic change of the reference coordinate adds a
\(\Lambda\)-independent function and does not change the derivative.

Canonical gauge couplings are different coordinates.  Passing to them
requires rescaling matter and gauge multiplets; the corresponding local
Berezinian is the Konishi-type anomaly derived below.  Thus the
one-loop-exact statement here is compatible with higher-loop beta functions
in canonical normalization.
\end{proof}

\section{Konishi-Type Rescaling Anomaly}

Let \(\Phi\) be a chiral superfield valued in a unitary representation \(R\)
of the gauge group.  The representation index used in this section is fixed
by
\[
  \operatorname{Tr}_{R}(XY)=T(R)\operatorname{tr}(XY),
  \qquad X,Y\in\mathfrak g,
\]
where \(\operatorname{tr}\) is the same invariant form used in the
Yang--Mills term
\(-\frac1{4g^2}\operatorname{tr}(F_{\mu\nu}F^{\mu\nu})\).

\begin{proposition}[Konishi rescaling Jacobian]
\label{prop:konishi-rescaling-jacobian}
Assume a gauge-covariant supersymmetric finite-mode regulator, or a
perturbative spectral regulator whose local heat-kernel coefficient agrees
with the finite-mode Berezinian limit, for the chiral superfield
\(\Phi\in R\).  Under the constant holomorphic field redefinition
\[
  \Phi\mapsto e^{\alpha}\Phi
\]
the gauge-field-dependent local part of the regulated super-Berezinian is
equivalent to adding the chiral local functional
\[
  \Delta S
  =
  \frac{T(R)\alpha}{8\pi^2}
  \int d^4x\,d^2\theta\,\operatorname{tr}(W^\alpha W_\alpha)
  +\text{c.c.}
\]
to the Wilsonian action, up to gauge-independent constants and local
\(D\)-terms removable by the coordinate choice.  For several chiral
superfields the right hand side is summed over their representations and
rescaling parameters.
\end{proposition}

\begin{proof}
The statement is a local Jacobian statement inside a specified regulator, not
a property of a formal infinite-dimensional measure.  First impose an
infrared auxiliary regulator so that the kinetic operator has a discrete
orthonormal basis.  The odd component field \(\psi_\alpha\) of \(\Phi\) is a
left Weyl fermion in \(R\).  Expanding it in a gauge-covariant finite set of
modes gives odd coordinates \(c_n\), and the ordered Berezinian density
\(\prod_n dc_n\) transforms by the inverse determinant of the finite matrix
representing the change of variables.  The scalar and auxiliary components
produce ordinary determinants.  Their gauge-dependent local contribution is
a \(D\)-term or a gauge-independent normalization in a supersymmetric
regulator; the chiral gauge-field-dependent term is fixed by the Weyl-fermion
Berezinian.

The finite-mode calculation is the chiral half of the Berezinian trace in
Chapter~\ref{ch:volume-ii-20}.  Replacing the sharp finite-mode projector by
the equivalent smooth gauge-covariant spectral regulator for the local
coefficient gives
\[
  \log J_{\alpha,M}^{\rm loc}
  =
  \alpha\,
  \operatorname{Tr}_{R,S_+}
  \left[
    \Pi_+\,\exp(-\mathcal D_A^2/M^2)
  \right]_{\rm loc}
  +\text{anti-chiral part}+O(M^{-1}),
\]
where \(S_+\) is the left Weyl spinor bundle and \(\mathcal D_A\) is the
Dirac operator coupled to the same connection.  The symbol
\([\cdots]_{\rm loc}\) means the part depending locally on the background
curvature after gauge-independent power divergences have been subtracted.
The chiral asymmetry of the heat kernel is
\[
  \lim_{M\to\infty}
  \operatorname{tr}_{S}
  \left[
    \gamma_5\exp(-\mathcal D_A^2/M^2)
  \right](x,x)
  =
  \frac{1}{32\pi^2}
  \epsilon^{\mu\nu\rho\sigma}
  \operatorname{Tr}_{R}(F_{\mu\nu}F_{\rho\sigma}),
\]
as in
\eqref{eq:heat-kernel-anomaly-density}.  Therefore the imaginary part of the
Jacobian for the phase subcase \(\alpha=\ii\beta\) is the regulated
Konishi-current anomaly of the left Weyl fermion in \(R\).  Using
\(\operatorname{Tr}_{R}(XY)=T(R)\operatorname{tr}(XY)\), its nontrivial
curvature dependence is proportional to \(T(R)\operatorname{tr}(F\wedge F)\).

It remains to identify the superspace representative of this local
functional.  The holomorphic part of the Jacobian must be a chiral,
gauge-invariant, local functional, linear in \(\alpha\), quadratic in the
background gauge field at leading order, and of engineering dimension three
in the chiral integrand.  In an \(\mathcal N=1\) gauge background the unique
such representative, modulo chiral total derivatives and \(D\)-terms, is
\[
  c\,\alpha\int d^4x\,d^2\theta\,
  \operatorname{tr}(W^\alpha W_\alpha).
\]
The coefficient \(c\) is determined by the component anomaly just computed.
With the superspace normalization of
Chapter~\ref{ch:susy-gauge-theory},
\[
  \frac{1}{16\pi\ii}\tau
  \int d^4x\,d^2\theta\,\operatorname{tr}(W^\alpha W_\alpha)
  +\text{c.c.}
\]
has component theta term
\(\frac{\theta}{32\pi^2}
\operatorname{tr}(F_{\mu\nu}\widetilde F^{\mu\nu})\) when
\(\operatorname{Re}\tau=\theta/(2\pi)\).  This fixes
\[
  c=\frac{T(R)}{8\pi^2}
\]
in the action-shift convention of the proposition.  The conjugate
anti-chiral Berezinian gives the displayed c.c. term.  Gauge-independent
power-divergent constants and local full-superspace terms change the
normalization of fields or the \(D\)-term coordinate and are not part of the
Konishi chiral anomaly.
\end{proof}

\begin{remark}[Meaning of the action-shift convention]
The sign of a Jacobian may be moved between the transformed density and the
action depending on whether one writes the change of variables as
\(\mathcal D\Phi=J\,\mathcal D\Phi'\) or
\(\mathcal D\Phi'=J^{-1}\mathcal D\Phi\), and on Lorentzian versus Euclidean
weight conventions.  Proposition~\ref{prop:konishi-rescaling-jacobian} fixes
\(\Delta S\) by the convention used in the Wilsonian coordinate relation
below: a holomorphic rescaling of matter fields shifts the holomorphic gauge
coordinate by the amount encoded in the displayed
\(\frac{T(R)\alpha}{8\pi^2}\int W^2\) term.  Ward identities obtained by
varying the same change of variables carry the corresponding term on the
opposite side if the Jacobian is kept in the measure instead of exponentiated
into the action.
\end{remark}

This anomaly explains why holomorphic Wilsonian couplings and canonically
normalized couplings obey different coordinate beta functions.  It is not a
claim that a manifestly supersymmetric nonperturbative regulator has been
constructed for every four-dimensional supersymmetric gauge theory; it is a
perturbative local Jacobian statement in schemes where the above regulated
Berezinian calculation is valid.

\section{NSVZ Coordinate Formula}

Fix the bilinear form \(\operatorname{tr}\) of
Chapter~\ref{ch:susy-gauge-theory}.  Define representation indices by
\[
  \operatorname{Tr}_{R_i}(XY)=T(R_i)\operatorname{tr}(XY),
  \qquad
  \operatorname{Tr}_{\mathrm{ad}}(XY)=C_2(G)\operatorname{tr}(XY).
\]
Let \(g_h\) be the holomorphic Wilsonian coupling and let \(g\) be the
canonically normalized gauge coupling in the same momentum-subtraction
scheme for the two-point function.  Let \(Z_i(\mu)\) be the matter
wave-function coordinate and define the anomalous-dimension convention
\[
  \gamma_i(\mu)=\mu\frac{d}{d\mu}\log Z_i(\mu).
\]

\begin{proposition}[Holomorphic-canonical gauge-coordinate relation]
\label{prop:nsvz-coordinate-relation-derived}
Assume a perturbative supersymmetric Wilsonian scheme in which:
\begin{enumerate}[label=(\roman*)]
\item the holomorphic gauge coordinate
\[
  X_h(\mu):=\frac{8\pi^2}{g_h(\mu)^2}
\]
is defined by the chiral gauge kinetic term in the convention of
Chapter~\ref{ch:susy-gauge-theory};
\item the canonical coordinate \(g(\mu)\) is defined by normalizing the
gauge-field two-point function so that the component kinetic term is
\(-\frac{1}{4g(\mu)^2}\operatorname{tr}(F_{\mu\nu}F^{\mu\nu})\);
\item the chiral matter kinetic terms are normalized by positive
wave-function coordinates \(Z_i(\mu)\);
\item the matter rescaling Jacobian is
Proposition~\ref{prop:konishi-rescaling-jacobian}, and the gauge-multiplet
rescaling Jacobian is computed in the same background-gauge BV regulator.
\end{enumerate}
Then, through the perturbative order to which the regulator and local
coordinates have been constructed,
\[
  \frac{8\pi^2}{g_h^2}
  =
  \frac{8\pi^2}{g^2}
  +
  C_2(G)\log g^2
  -
  \sum_i T(R_i)\log Z_i
  +
  \kappa ,
\]
where \(\kappa\) is independent of \(\mu\) and records finite choices of
reference normalization for the local coordinates.
\end{proposition}

\begin{proof}
Write the chiral gauge-coordinate part of the Wilsonian action as
\[
  S_{\rm gauge}^{F}
  =
  \frac{X_h}{16\pi^2}
  \int d^4x\,d^2\theta\,
  \operatorname{tr}(W^\alpha W_\alpha)
  +\text{theta-angle part}+\text{c.c.}
\]
This equation fixes what is meant by \(X_h=8\pi^2/g_h^2\).  With the same
normalization, Proposition~\ref{prop:konishi-rescaling-jacobian} says that a
holomorphic rescaling \(\Phi_i\mapsto e^{\alpha_i}\Phi_i\) shifts this real
coordinate by
\begin{equation}
  \Delta_i X_h=2T(R_i)\alpha_i .
  \label{eq:konishi-shift-Xh}
\end{equation}
Indeed the chiral action coefficient changes by
\(\Delta_i X_h/(16\pi^2)\), while the Jacobian coefficient in
Proposition~\ref{prop:konishi-rescaling-jacobian} is
\(T(R_i)\alpha_i/(8\pi^2)\).

To pass from the Wilsonian fields with matter kinetic coordinate \(Z_i\) to
canonical matter fields, set
\[
  \Phi_i^{\rm hol}=Z_i^{-1/2}\Phi_i^{\rm can}.
\]
Thus \(\alpha_i=-\frac12\log Z_i\).  Substitution in
\eqref{eq:konishi-shift-Xh} gives the matter contribution
\begin{equation}
  \Delta_{\rm matter}X_h
  =
  -\sum_iT(R_i)\log Z_i .
  \label{eq:matter-rescaling-shift-Xh}
\end{equation}
This is the part of the relation that is often called the Konishi anomaly.
It is a statement about the local Berezinian of the regulated change of
variables, not about an unregulated functional measure.

The vector multiplet must also be rescaled when passing from holomorphic
normalization to the canonical gauge coordinate.  In a background-field BV
gauge fixing, the rescaling is performed on the whole quantum gauge complex:
the vector superfield, ghosts, antighosts, auxiliary fields, and antifields
are rescaled so that the BV pairing and gauge-fixed quadratic form are kept
in canonical dual normalization.  Let
\[
  V^{\rm hol}=g\,V^{\rm can},
  \qquad
  \rho:=\log g=\frac12\log g^2 .
\]
The regulated super-Berezinian of this adjoint BV complex has the chiral
local part
\[
  \Delta_{\rm vec}S
  =
  \frac{C_2(G)\rho}{8\pi^2}
  \int d^4x\,d^2\theta\,
  \operatorname{tr}(W^\alpha W_\alpha)
  +\text{c.c.}
\]
Equivalently,
\begin{equation}
  \Delta_{\rm vec}X_h
  =
  2C_2(G)\rho
  =
  C_2(G)\log g^2 .
  \label{eq:gauge-multiplet-rescaling-shift-Xh}
\end{equation}
Here \(C_2(G)\) is defined by
\(\operatorname{Tr}_{\rm ad}(XY)=C_2(G)\operatorname{tr}(XY)\).  The
coefficient is obtained from the same heat-kernel coefficient as in the
matter calculation, but with the supertrace over the gauge-fixed adjoint BV
complex.  The longitudinal vector modes and chiral Faddeev--Popov/Nielsen--
Kallosh ghost modes cancel except for the adjoint chiral index contribution
shown above.  Performing the rescaling on only \(A_\mu\) and ignoring this
complex would not define a gauge-consistent Jacobian.

Classically, after the vector and matter variables have been brought to the
canonical coordinate chart, the real gauge kinetic coordinate is
\[
  X_c:=\frac{8\pi^2}{g^2}.
\]
Adding the two anomalous local shifts
\eqref{eq:matter-rescaling-shift-Xh} and
\eqref{eq:gauge-multiplet-rescaling-shift-Xh}, and allowing a finite
\(\mu\)-independent normalization difference between the two declared
coordinate charts, gives
\[
  X_h
  =
  X_c
  +
  C_2(G)\log g^2
  -
  \sum_iT(R_i)\log Z_i
  +
  \kappa .
  \]
This is the displayed relation.  The constant \(\kappa\) changes under finite
renormalizations such as \(Z_i\mapsto c_iZ_i\), \(g\mapsto g+O(g^3)\) at a
chosen reference scale, or a finite local gauge kinetic counterterm; it has
zero \(\mu\)-derivative in the fixed scheme and therefore drops out of the
beta-function identity.
\end{proof}

\begin{proposition}[NSVZ beta function as a coordinate identity]
\label{prop:nsvz-beta-coordinate-identity}
Assume the hypotheses of
Proposition~\ref{prop:nsvz-coordinate-relation-derived}, and assume the
holomorphic Wilsonian coordinate has the one-loop derivative
\[
  \mu\frac{d}{d\mu}\left(\frac{8\pi^2}{g_h^2}\right)
  =
  3C_2(G)-\sum_iT(R_i).
  \label{eq:holomorphic-coordinate-one-loop-derivative}
\]
Then, in the canonical coupling coordinate \(g\), with
\[
  \gamma_i=\mu\,\frac{d}{d\mu}\log Z_i ,
\]
\[
  \beta_g
  =
  -\frac{g^3}{16\pi^2}\,
  \frac{
    3C_2(G)-\sum_iT(R_i)\bigl(1-\gamma_i\bigr)
  }{
    1-\frac{C_2(G)g^2}{8\pi^2}
  } .
  \tag{NSVZ}
  \label{eq:nsvz-coordinate-beta-function}
\]
\end{proposition}

\begin{proof}
Differentiate the coordinate relation of
Proposition~\ref{prop:nsvz-coordinate-relation-derived}.  Since
\[
  \mu\frac{d}{d\mu}\left(\frac{8\pi^2}{g^2}\right)
  =
  -\frac{16\pi^2}{g^3}\beta_g,
  \qquad
  \mu\frac{d}{d\mu}\log g^2=\frac{2\beta_g}{g},
\]
and \(\mu\,d\kappa/d\mu=0\), one obtains
\[
  \left(
    -\frac{16\pi^2}{g^3}
    +
    \frac{2C_2(G)}{g}
  \right)\beta_g
  -
  \sum_iT(R_i)\gamma_i
  =
  3C_2(G)-\sum_iT(R_i).
\]
Move the anomalous-dimension term to the right:
\[
  \left(
    \frac{16\pi^2}{g^3}
    -
    \frac{2C_2(G)}{g}
  \right)\beta_g
  =
  -
  \left(
    3C_2(G)-\sum_iT(R_i)(1-\gamma_i)
  \right).
\]
Factoring the left hand side as
\[
  \frac{16\pi^2}{g^3}
  \left(
    1-\frac{C_2(G)g^2}{8\pi^2}
  \right)\beta_g
\]
and dividing gives \eqref{eq:nsvz-coordinate-beta-function}.  Thus the
displayed beta function is a coordinate statement derived from a holomorphic
Wilsonian coupling, the rescaling anomaly, and a declared canonical
normalization prescription.  A finite redefinition of the canonical coupling
changes the denominator by ordinary reparametrization of the vector field
\(\beta_g\partial_g\).
\end{proof}

\section{Status Boundary}

The perturbative proposition above is a Wilsonian statement under explicit
regularity hypotheses.  Nonperturbative superpotentials, instanton effects,
gaugino condensation, and strong-coupling moduli-space constraints are not
contradictions: they are additional semiclassical or nonperturbative
contributions to Wilsonian chiral coordinates.  Each such contribution must
be derived from a declared integration cycle, zero-mode measure, anomaly
matching, and infrared effective description.

\begin{openproblem}[Rigorous supersymmetric Wilsonian holomorphy]
\label{op:rigorous-susy-holomorphy}
Construct a regulator and BV Wilsonian scheme in which the chiral local
cohomology, supergraph measure, rescaling anomaly, and holomorphic gauge
coordinate are all defined at finite cutoff, and prove the perturbative
nonrenormalization statements in that scheme.
\end{openproblem}
