\chapter{Two-Dimensional Supersymmetric Models}
\label{ch:susy-two-dimensional-lg_sigma_glsm}

\ConventionsMixed

This chapter begins the lower-dimensional supersymmetric examples.  The main
objects are \(2D\) \(\mathcal N=(2,2)\) Landau--Ginzburg models, sigma
models, and gauged linear sigma models.  Each is treated as a QFT presentation with
declared fields, symmetries, operator coordinates, and infrared questions.
Flavor and \(R\)-symmetry data in this chapter are therefore not silently
interpreted as ordinary local continuous order parameters.  In
\(1+1\)-dimensional unitary local QFT the Coleman obstruction of
Section~\ref{sec:low-dimensional-continuous-ssb-no-go} is the relevant
default; protected chiral rings, indices, anomalies, twists, and topological
sectors are different infrared data.

\section{The N=(2,2) Algebra}

In Lorentzian signature let \(P_\pm=P_0\pm P_1\).  The supercharges
\[
  Q_\pm,\quad \bar Q_\pm
\]
obey
\[
  \{Q_+,\bar Q_+\}=2P_+,\qquad
  \{Q_-,\bar Q_-\}=2P_-,
\]
with all other anticommutators vanishing in the absence of central charges.
An \(R\)-symmetry assignment consists of vector and axial charges
\[
  F_V,\qquad F_A,
\]
whose action on supercharges is part of the model specification.  A BPS state with
central charge \(Z\) obeys an inequality obtained by positivity of
\((Q+\ee^{\ii\alpha}\bar Q)^\dagger(Q+\ee^{\ii\alpha}\bar Q)\); the saturated
states are kernels of a corresponding linear combination of supercharges.

\section{Landau--Ginzburg Presentations}

\begin{definition}[Two-dimensional Landau--Ginzburg presentation]
\label{def:two-dimensional-lg-presentation}
A polynomial \(2D\) \(\mathcal N=(2,2)\) Landau--Ginzburg presentation in this
chapter is specified by:
\begin{enumerate}
\item a Euclidean two-manifold \(\Sigma\), initially a smooth oriented
surface with a chosen spin structure when fermions are kept explicitly;
\item chiral multiplet coordinates
\[
  \Phi^i=(\phi^i,\psi^i_+,\psi^i_-,F^i),
  \qquad i=1,\ldots,n,
\]
where \(\phi^i:\Sigma\to\C\) are scalar field variables, \(\psi^i_\pm\) are
fermionic field variables of right- and left-moving chirality, and \(F^i\)
are auxiliary even field variables;
\item a K\"ahler potential \(K\) whose Hessian
\[
  g_{i\bar j}(\phi,\bar\phi)=\partial_i\partial_{\bar j}K(\phi,\bar\phi)
\]
is positive on the field chart used in the perturbative construction;
\item a holomorphic polynomial superpotential
\[
  W\in\C[\phi^1,\ldots,\phi^n].
\]
\end{enumerate}
This is a field-variable presentation.  A continuum QFT requires a
regulator, a supersymmetric Ward identity, and a Hilbert-space or Euclidean
reconstruction; none of these is inferred from the polynomial \(W\) alone.
\end{definition}

Eliminating the auxiliary fields is a finite-dimensional calculation at each
point of \(\Sigma\).  In the normalization where the auxiliary part of the
Euclidean Lagrangian is
\[
  g_{i\bar j}F^i\bar F^{\bar j}
  +F^i\partial_iW+\bar F^{\bar j}\partial_{\bar j}\overline W ,
\]
the algebraic equation of motion is
\[
  g_{i\bar j}\bar F^{\bar j}+\partial_iW=0 .
\]
Substituting this equation gives the bosonic potential
\[
  V(\phi)=g^{i\bar j}\partial_iW(\phi)\,
  \overline{\partial_jW(\phi)}.
\]
Thus supersymmetric classical vacua in the affine field chart solve
\[
  \partial_iW(\phi)=0,\qquad i=1,\ldots,n.
\]

\begin{definition}[Quasihomogeneous isolated Landau--Ginzburg polynomial]
\label{def:vii-quasihomogeneous-lg-polynomial}
Let \(q_i\in\mathbb Q_{>0}\).  A polynomial
\[
  W\in\C[\phi^1,\ldots,\phi^n]
\]
is quasihomogeneous of degree \(1\) with weights \(q_i\) when
\begin{equation}
  W(\lambda^{q_1}\phi^1,\ldots,\lambda^{q_n}\phi^n)
  =
  \lambda W(\phi^1,\ldots,\phi^n)
  \qquad
  \lambda\in\C^\times .
  \label{eq:vii-lg-quasihomogeneous-condition}
\end{equation}
It is isolated when the common zero locus of
\[
  \partial_1W,\ldots,\partial_nW
\]
in \(\C^n\) is the single point \(0\).  The algebraic chiral coordinate ring
attached to the polynomial is the Jacobi algebra
\begin{equation}
  \operatorname{Jac}(W)
  =
  \C[\phi^1,\ldots,\phi^n]/
  (\partial_1W,\ldots,\partial_nW).
  \label{eq:vii-lg-jacobi-ring}
\end{equation}
\end{definition}

The term ``algebraic chiral coordinate ring'' is deliberately narrower than
``the infrared chiral ring of the continuum QFT''.  It is the finite
algebraic object forced by the \(F\)-term equations and by the protected
cohomological argument below.  The statement that a regulated Landau--Ginzburg
model flows to a compact \(\mathcal N=2\) superconformal field theory with
this ring as its chiral-primary ring is an additional infrared construction
problem.

\paragraph{Jacobian quotient from the \(B\)-type differential.}
Let \(A=\C[\phi^1,\ldots,\phi^n]\) and let \(I_W\subset A\) be the ideal
generated by \(\partial_iW\).  Assume that the regulated \(B\)-type local
operator complex contains odd local variables \(\eta_i\), \(i=1,\ldots,n\),
and a scalar odd differential \(Q_B\) satisfying
\begin{equation}
  Q_B\phi^i=0,\qquad
  Q_B\eta_i=\partial_iW(\phi),
  \qquad
  Q_B^2=0
  \label{eq:lg-btype-local-differential}
\end{equation}
on polynomial local coordinates.  Then the image of \(Q_B\) inside the
polynomial subalgebra \(A\) is exactly \(I_W\).  Hence the polynomial
\(Q_B\)-cohomology in degree zero is canonically
\[
  H^0_{Q_B}\bigl(A\otimes\Lambda(\eta_1,\ldots,\eta_n)\bigr)
  \cong
  \operatorname{Jac}(W).
\]

Every polynomial \(f(\phi)\in A\) is \(Q_B\)-closed because
\(Q_B\phi^i=0\) and \(Q_B\) is a derivation.  Conversely, the \(Q_B\)-exact
polynomials are generated by applying \(Q_B\) to degree-one expressions
\[
  \sum_i h^i(\phi)\eta_i,\qquad h^i\in A.
\]
Using \eqref{eq:lg-btype-local-differential},
\[
  Q_B\left(\sum_i h^i(\phi)\eta_i\right)
  =
  \sum_i h^i(\phi)\partial_iW(\phi),
\]
because \(Q_Bh^i(\phi)=0\).  Thus
\(\operatorname{im}Q_B\cap A=I_W\), and quotienting the closed polynomial
operators by exact polynomial operators gives \eqref{eq:vii-lg-jacobi-ring}.

\paragraph{Quasihomogeneous charge and central-charge test.}
Let \(W\) be a quasihomogeneous isolated polynomial in the sense of
Definition~\ref{def:vii-quasihomogeneous-lg-polynomial}.  Assign the monomial
\[
  \phi^a=(\phi^1)^{a_1}\cdots(\phi^n)^{a_n}
\]
the charge
\begin{equation}
  Q_R(\phi^a)=\sum_{i=1}^n a_iq_i .
  \label{eq:vii-lg-monomial-charge}
\end{equation}
Then every monomial appearing in \(W\) has charge \(1\), while every monomial
appearing in \(\partial_iW\) has charge \(1-q_i\).  If, in addition,
\begin{equation}
  W=\sum_{i=1}^n(\phi^i)^{d_i},
  \qquad d_i\geq2,
  \label{eq:fermat-lg-polynomial}
\end{equation}
then \(q_i=1/d_i\), the Jacobi algebra has basis
\begin{equation}
  \prod_{i=1}^n(\phi^i)^{a_i},
  \qquad
  0\leq a_i\leq d_i-2,
  \label{eq:fermat-lg-jacobi-basis}
\end{equation}
and dimension
\begin{equation}
  \dim_\C\operatorname{Jac}(W)=\prod_{i=1}^n(d_i-1).
  \label{eq:fermat-lg-jacobi-dimension}
\end{equation}
The protected \(U(1)_R\)-anomaly prediction for a compact unitary infrared
\(\mathcal N=2\) superconformal fixed point is
\begin{equation}
  c_{\rm LG}=3\sum_{i=1}^n(1-2q_i).
  \label{eq:vii-lg-central-charge-prediction}
\end{equation}
In this chapter \eqref{eq:vii-lg-central-charge-prediction} is used as a
necessary protected-data test for any proposed compact unitary infrared fixed
point with these chiral fields and this \(U(1)_R\) current.

Equation~\eqref{eq:vii-lg-quasihomogeneous-condition} says precisely that
every monomial in \(W\) has total charge \(1\).  Differentiating a monomial
with respect to \(\phi^i\) lowers its charge by \(q_i\), hence the terms in
\(\partial_iW\) have charge \(1-q_i\).

For the Fermat polynomial \eqref{eq:fermat-lg-polynomial},
quasihomogeneity gives \(d_iq_i=1\), so \(q_i=1/d_i\).  The Jacobi ideal is
\[
  ((\phi^1)^{d_1-1},\ldots,(\phi^n)^{d_n-1}),
\]
up to multiplication by the nonzero constants \(d_i\).  Reduction modulo this
monomial ideal removes every factor with exponent \(a_i\geq d_i-1\), and the
remaining monomials \eqref{eq:fermat-lg-jacobi-basis} are linearly
independent in the quotient.  Counting the allowed exponents gives
\eqref{eq:fermat-lg-jacobi-dimension}.

The final formula is the protected anomaly calculation used by the
\(\mathcal N=2\) superconformal-algebra interface in
Chapter~\ref{ch:two-dimensional-superconformal-algebras}.  The present
paragraph records its algebraic inputs and status: once a continuum
infrared fixed point with these chiral fields and \(U(1)_R\) current has
been constructed, the anomaly supplies \eqref{eq:vii-lg-central-charge-prediction}.

This is therefore an algebraic and anomaly-matching datum.  The existence of
the infrared fixed point and the convergence of a Landau--Ginzburg regulator
to it are separate construction and RG questions.

\paragraph{Wilsonian superpotential coordinate for Fermat LG models.}
The polynomial \(W\) is treated here as part of the Wilsonian \(F\)-term
coordinate of a declared regulated theory.  The following
finite argument is the two-dimensional version of the holomorphic-coordinate
logic of Chapter~\ref{ch:susy-nonrenormalization-holomorphy}.

Fix
\[
  W_{\rm UV}(\Phi;g)=\sum_{i=1}^n g_i(\Phi^i)^{d_i},
  \qquad d_i\geq2,
\]
and assume a Wilsonian scheme at scale \(\Lambda\) with the following
properties:
\begin{enumerate}
\item the chiral superspace coordinate is preserved, so corrections to the
  superpotential are holomorphic functions of the chiral fields and of the
  background chiral couplings \(g_i\);
\item the patch is regular at \(g_i=0\), so the perturbative Wilsonian
  expansion contains nonnegative powers of the \(g_i\);
\item the scheme preserves the \(n\) flavor spurion symmetries
  \[
    \Phi^i\mapsto \lambda_i\Phi^i,\qquad
    g_i\mapsto \lambda_i^{-d_i}g_i,
    \qquad \lambda_i\in\C^\times,
  \]
  together with the quasihomogeneous \(R\)-charge convention
  \[
    R(\Phi^i)=\frac1{d_i},\qquad R(g_i)=0,\qquad R(W)=1 .
  \]
\end{enumerate}
Then every regular monomial allowed in the Wilsonian superpotential is one of
the original monomials \(g_i(\Phi^i)^{d_i}\), up to its coefficient.  Indeed,
write a candidate monomial as
\[
  \prod_{i=1}^n g_i^{m_i}(\Phi^i)^{a_i},
  \qquad m_i,a_i\in\mathbb Z_{\geq0}.
\]
The \(i\)-th spurion symmetry requires
\begin{equation}
  a_i-d_i m_i=0
  \qquad\text{for every }i,
  \label{eq:two-d-lg-spurion-neutrality}
\end{equation}
and the chiral \(R\)-charge condition requires
\begin{equation}
  \sum_{i=1}^n\frac{a_i}{d_i}=1 .
  \label{eq:two-d-lg-r-charge-selection}
\end{equation}
Substituting \eqref{eq:two-d-lg-spurion-neutrality} into
\eqref{eq:two-d-lg-r-charge-selection} gives
\[
  \sum_{i=1}^n m_i=1.
\]
Since the \(m_i\) are nonnegative integers, exactly one \(m_j\) is \(1\) and
all other \(m_i\) vanish.  Then \(a_j=d_j\) and \(a_i=0\) for \(i\neq j\).
Thus the holomorphic selection rule alone permits only
\[
  g_j(\Phi^j)^{d_j},
  \qquad j=1,\ldots,n.
\]

The scope of the argument is the Wilsonian coordinate at the declared scale:
after a regulator with the stated chiral and spurion Ward identities has
been supplied, perturbative local corrections preserve the Fermat
superpotential coordinate up to finite redefinition of the displayed
couplings.  Existence of the continuum infrared fixed point, convergence of
the RG flow, and control of any allowed nonperturbative sectors are separate
analytic inputs.  The K\"ahler potential and other \(D\)-term coordinates
may still flow.

The same protected-data interface controls the finite \(\chi_y\) shadow of
the elliptic genus in
Section~\ref{sec:elliptic-genus-spectral-flow-covariance}.  If the infrared
fixed point exists and the Jacobi algebra classes become NS chiral primaries,
then spectral flow by \(\eta=-1/2\) sends a monomial of charge \(Q_R\) to a
left Ramond ground-state contribution of charge
\begin{equation}
  Q_R-\frac{c_{\rm LG}}6 .
  \label{eq:vii-lg-ramond-charge-shift}
\end{equation}
Thus the finite polynomial \(\sum_{[O]\in\operatorname{Jac}(W)}
y^{Q_R(O)-c_{\rm LG}/6}\) is a necessary protected check on the proposed
SCFT.  The full elliptic genus requires the compact Ramond trace and modular
sewing hypotheses stated in the CFT volume.

\begin{example}[\(A\)-series and quintic polynomial checks]
\label{ex:vii-lg-a-series-quintic-checks}
For \(W=X^{k+2}\), \(k\geq0\), one has \(q=1/(k+2)\),
\[
  \operatorname{Jac}(W)=\C[X]/(X^{k+1}),
  \qquad
  c_{\rm LG}=3\left(1-\frac{2}{k+2}\right)=\frac{3k}{k+2}.
\]
For
\[
  W=X_1^5+\cdots+X_5^5
\]
the weights are \(q_i=1/5\), the algebraic Jacobi dimension is
\[
  4^5=1024,
\]
and the protected central-charge test gives
\[
  c_{\rm LG}=3\cdot5\left(1-\frac25\right)=9.
\]
This is an intrinsic Landau--Ginzburg charge and Jacobi-algebra calculation.
\end{example}

\section{Sigma-Model Datum}

A \(2D\) \(\mathcal N=(2,2)\) nonlinear sigma model is specified by a K\"ahler
target \(X\), metric \(g_{i\bar j}\), \(B\)-field \(B\), and possibly a
superpotential \(W\).  The bosonic Euclidean action is
\[
  S_{\mathrm{bos}}
  =
  \frac1{4\pi\alpha'}
  \int_\Sigma
  \dd^2x\,
  g_{i\bar j}(\phi)\partial_\mu\phi^i\partial_\mu\bar\phi^{\bar j}
  +
  \frac{\ii}{2\pi\alpha'}
  \int_\Sigma \phi^*B .
\]
The large-volume perturbative beta tensor begins with the Ricci form:
\[
  \mu\frac{\dd g_{i\bar j}}{\dd\mu}
  =
  \frac1{2\pi}R_{i\bar j}+\cdots .
\]
Thus Ricci-flatness is the first condition for a conformal infrared candidate
in a large-volume chart.  The full QFT claim includes instanton corrections,
operator renormalization, and a continuum limit; these are additional data,
separate from the local tensor equation.

\section{A And B Twisting Data}

The two cohomological twists of a \(2D\) \(\mathcal N=(2,2)\) theory are
intrinsic operations on the two-dimensional QFT.  They require nonanomalous
\(R\)-symmetries and a regulator preserving the selected scalar supercharge.
The definition uses the two-dimensional rotation symmetry, the selected
\(R\)-current, and the supersymmetry algebra.

\begin{definition}[Two-dimensional twist convention]
\label{def:two-dimensional-ab-twist-convention}
Let \(s\) denote the Lorentz spin for \(U(1)_{\rm rot}\).  In this chapter
the supercharges have charges
\begin{equation}
\begin{array}{c|rrrr}
  & Q_+ & \bar Q_+ & Q_- & \bar Q_-\\ \hline
  s & \frac12 & \frac12 & -\frac12 & -\frac12\\
  F_V & 1 & -1 & 1 & -1\\
  F_A & 1 & -1 & -1 & 1 .
\end{array}
\label{eq:two-dimensional-supercharge-charge-table}
\end{equation}
The A twist uses
\begin{equation}
  s_A=s+\frac12F_V ,
  \label{eq:a-twist-spin-shift}
\end{equation}
and the B twist uses
\begin{equation}
  s_B=s+\frac12F_A .
  \label{eq:b-twist-spin-shift}
\end{equation}
\end{definition}

\begin{remark}[Twist bookkeeping: scalar supercharges]
\label{rem:two-dimensional-ab-twist-scalar-supercharges}
Assume the charge convention of
Definition~\ref{def:two-dimensional-ab-twist-convention} and assume that
there are no central charges in the local-sector supersymmetry algebra.
Then:
\begin{equation}
  Q_A=\bar Q_+ + Q_-
  \label{eq:a-twist-scalar-supercharge}
\end{equation}
is scalar for the A-twisted rotation action and satisfies \(Q_A^2=0\) on
gauge-invariant local observables, while
\begin{equation}
  Q_B=\bar Q_+ + \bar Q_-
  \label{eq:b-twist-scalar-supercharge}
\end{equation}
is scalar for the B-twisted rotation action and satisfies \(Q_B^2=0\) on
gauge-invariant local observables.

Using \eqref{eq:a-twist-spin-shift}, the A-twisted spins are
\[
\begin{array}{c|rrrr}
  & Q_+ & \bar Q_+ & Q_- & \bar Q_-\\ \hline
  s_A & 1 & 0 & 0 & -1 .
\end{array}
\]
Thus \(\bar Q_+\) and \(Q_-\) are scalars, while the other two supercharges
are not.  Using \eqref{eq:b-twist-spin-shift}, the B-twisted spins are
\[
\begin{array}{c|rrrr}
  & Q_+ & \bar Q_+ & Q_- & \bar Q_-\\ \hline
  s_B & 1 & 0 & -1 & 0 .
\end{array}
\]
Thus \(\bar Q_+\) and \(\bar Q_-\) are scalars for the B twist.  The
untwisted local algebra has only
\[
  \{Q_+,\bar Q_+\}=2P_+,\qquad
  \{Q_-,\bar Q_-\}=2P_-
\]
among these generators.  Therefore
\[
  Q_A^2=\frac12\{\bar Q_+ + Q_-,\bar Q_+ + Q_-\}=0,
  \qquad
  Q_B^2=\frac12\{\bar Q_+ + \bar Q_-,\bar Q_+ + \bar Q_-\}=0
\]
in the central-charge-free local sector.  If gauge transformations appear in
an off-shell realization, the same statement is made after passing to
gauge-invariant observables.
\end{remark}

\section{Gauged Linear Sigma-Model Presentations}

\begin{definition}[Abelian \(2D\) \(\mathcal N=(2,2)\) GLSM presentation]
\label{def:abelian-2d-glsm-presentation}
An abelian GLSM presentation in this chapter is specified by:
\begin{enumerate}
\item gauge group \(G=U(1)\);
\item chiral multiplets \(\Phi_i\), \(i=1,\ldots,m\), with integer charges
  \(Q_i\in\mathbb Z\);
\item a \(U(1)\)-invariant holomorphic superpotential
  \(W\in\C[\Phi_1,\ldots,\Phi_m]^{U(1)}\);
\item an FI parameter \(r\), a theta angle \(\theta\), and the two associated
  holomorphic coordinates
\[
  \tau=\frac{\theta}{2\pi}+\ii r,
  \qquad
  T=2\pi\ii\,\tau=-2\pi r+\ii\theta,
  \qquad
  q=\ee^T ,
\]
  with \(\theta\in\mathbb R/2\pi\mathbb Z\), so
  \(\tau\sim\tau+1\), \(T\sim T+2\pi\ii\), and \(q\) is single-valued;
\item a gauge coupling \(e>0\) and a specified ultraviolet regulator for the
  gauge theory.
\end{enumerate}
The classical scalar potential contains
\begin{equation}
  \frac{e^2}{2}\left(\sum_{i=1}^m Q_i|\phi_i|^2-r\right)^2
  +\sum_i|\partial_iW|^2
  +|\sigma|^2\sum_i Q_i^2|\phi_i|^2 ,
  \label{eq:abelian-glsm-classical-potential}
\end{equation}
where \(\sigma\) is the complex scalar in the vector multiplet.  Classical
Higgs vacua with at least one charged scalar nonzero therefore solve the
\(F\)-term equations, the \(D\)-term equation, and then are quotiented by
\(U(1)\).
\end{definition}

\begin{remark}[Compact FI-theta normalization]
\label{rem:glsm-compact-fi-theta-normalization}
The coordinate used in exponentiated vortex, mirror, and Coulomb formulas in
this chapter is the logarithmic coordinate \(T\), not the period-one coordinate
\(\tau\).  The convention is fixed at the component level as follows.  The
unit-charge generator is normalized by
\[
  D_\mu\phi_i=(\partial_\mu+\ii Q_iA_\mu)\phi_i,\qquad
  F=dA,\qquad
  k=\frac1{2\pi}\int_\Sigma F\in\mathbb Z
\]
on an oriented Euclidean worldsheet.  The Euclidean auxiliary-field contour is
chosen so that eliminating \(D\) gives the positive term displayed in
\eqref{eq:abelian-glsm-classical-potential}.  With this normalization the
bosonic action contains the topological theta term
\[
  S_{E,\theta}=-\ii\theta k .
\]
In a positive vortex sector the Bogomolny completion of the gauge and
\(D\)-term energy supplies \(2\pi r k\), so the classical topological weight
of a flux-\(k\) saddle is
\begin{equation}
  \exp[-S_{E,{\rm top}}(k)]
  =
  \exp\!\left[(-2\pi r+\ii\theta)k\right]
  =
  q^k,
  \qquad q=\ee^T .
  \label{eq:glsm-compact-flux-topological-weight}
\end{equation}
For the opposite orientation or anti-vortex convention the inverse fugacity is
used, but the period test is the same: \(\theta\mapsto\theta+2\pi\) sends
\(T\mapsto T+2\pi\ii\) and leaves \(q^k\) unchanged for every integral flux.
The expression \(\ee^\tau\) is therefore not a physical fugacity in this
normalization.
\end{remark}

\begin{remark}[Twisted \(F\)-component normalization]
\label{rem:glsm-twisted-f-component-normalization}
The same compact convention fixes the superspace-to-component normalization used
below.  We use Euclidean light-cone derivatives with
\[
  \{D_+,\overline D_+\}=2\ii\partial_+,\qquad
  \{D_-,\overline D_-\}=2\ii\partial_-,
  \qquad
  \hbox{all other anticommutators zero}.
\]
The twisted chiral measure is ordered so that, for
\(\widetilde\theta=(\theta^-,\overline\theta^+)\),
\[
  \int d^2\widetilde\theta\,
  \theta^-\overline\theta^+=1 .
\]
In Wess--Zumino gauge, suppressing gaugino terms and derivative completions not
needed for the constant-background projection,
\[
  V_{\rm WZ}
  =
  \theta^-\overline\theta^- A_-
  +\theta^+\overline\theta^+ A_+
  -\theta^-\overline\theta^+\sigma
  -\theta^+\overline\theta^-\overline\sigma
  +\theta^-\theta^+\overline\theta^-\overline\theta^+D+\cdots ,
\]
with \(F_{12}=\partial_1A_2-\partial_2A_1\) and orientation
\(\dd x^1\wedge\dd x^2>0\).  The field strength used in this chapter is
\begin{equation}
  \Sigma=\overline D_+D_-V,
  \qquad
  \Sigma
  =
  \sigma+\cdots+
  \theta^-\overline\theta^+\bigl(D+\ii F_{12}\bigr)+\cdots .
  \label{eq:glsm-sigma-component-normalization}
\end{equation}
Thus the top component paired with a background variation of \(\Sigma\) is
\(\delta D+\ii\delta F_{12}\).  The normalized twisted-\(F\) component
projection is
\begin{equation}
  \delta{\mathcal I}_{\widetilde F}[\widetilde W]
  =
  \frac1{2\pi}
  \int d^2x\,
  \left[
    \Re\left(
      \frac{\partial\widetilde W}{\partial\Sigma}\right)\delta D
    +\ii\,
    \Im\left(
      \frac{\partial\widetilde W}{\partial\Sigma}\right)
    \delta F_{12}
  \right],
  \label{eq:glsm-twisted-f-component-pairing}
\end{equation}
with the c.c. term included in the displayed real and phase parts.  When
\(\widetilde W_{\rm FI}=-T\Sigma\),
\[
  \frac{\partial\widetilde W_{\rm FI}}{\partial\Sigma}
  =
  -T=2\pi r-\ii\theta ,
\]
so the Euclidean FI component action contains
\[
  S_{E,{\rm FI}}^{\rm lin}
  =
  \int d^2x\, rD
  -\frac{\ii\theta}{2\pi}\int F .
\]
Together with the \(D\)-contour and the Bogomolny completion this is the
component origin of the compact flux weight
\eqref{eq:glsm-compact-flux-topological-weight}.  Likewise, the local
dualization term
\(-Q\Sigma Y\) gives a dynamical theta angle: shifting
\(Y\mapsto Y+2\pi\ii\) changes the \(F_{12}\) phase by an integral multiple of
\(2\pi\ii k\) and hence leaves the path integral invariant.  The factor
\(1/(2\pi)\) in
\eqref{eq:glsm-twisted-f-component-pairing} is therefore fixed before the
Coulomb determinant is evaluated; it is not a convention chosen after the
momentum integral.
\end{remark}

The axial \(R\)-symmetry anomaly coefficient of this abelian presentation is
\begin{equation}
  \mathcal A_A=\sum_{i=1}^m Q_i.
  \label{eq:abelian-glsm-axial-anomaly}
\end{equation}
This coefficient is a finite charge sum: each charged chiral multiplet
contributes its charge to the axial rotation of the fermion measure.  When
\(\mathcal A_A=0\), the one-loop axial shift of the FI coordinate vanishes.
This is a necessary condition for an axial \(R\)-symmetry in an infrared
superconformal candidate.  A conformality claim additionally requires a
construction of the infrared continuum theory and its stress tensor.

The vector and axial \(R\)-symmetries have different anomaly requirements.
The A twist is available only when \(F_V\) is a genuine symmetry of the
regulated theory.  The B twist is available only when \(F_A\) is a genuine
symmetry.  In the abelian GLSM presentation of
Definition~\ref{def:abelian-2d-glsm-presentation}, the axial anomaly obstruction is
the charge sum \eqref{eq:abelian-glsm-axial-anomaly}.  Therefore the
hypersurface GLSM with charges \((1,\ldots,1,-d)\) has an axial B-twist
candidate precisely when \(N=d\).

\section{B-Twisted Landau--Ginzburg Sector}

The B-twisted Landau--Ginzburg model uses the scalar supercharge
\eqref{eq:b-twist-scalar-supercharge}.  The Jacobian-quotient calculation
above gives the first
protected observable: polynomial local operators map to
\(\operatorname{Jac}(W)\).  The next finite-dimensional check is the residue
pairing at nondegenerate critical points.

\paragraph{Morse residue pairing on the LG Jacobi algebra.}
Let \(W\in\C[\phi^1,\ldots,\phi^n]\) have a finite critical set
\[
  \operatorname{Crit}(W)=\{p\in\C^n:\partial_iW(p)=0
  \text{ for all }i\}.
\]
Assume every critical point is nondegenerate:
\[
  H_W(p)=\det(\partial_i\partial_jW)(p)\neq0 .
\]
Define
\begin{equation}
  \operatorname{Res}_W(f)
  =
  \sum_{p\in\operatorname{Crit}(W)}
  \frac{f(p)}{H_W(p)},
  \qquad f\in\C[\phi^1,\ldots,\phi^n].
  \label{eq:lg-morse-residue-functional}
\end{equation}
Then \(\operatorname{Res}_W\) kills the ideal
\((\partial_1W,\ldots,\partial_nW)\), so it descends to
\(\operatorname{Jac}(W)\).  The bilinear form
\begin{equation}
  \langle [f],[g]\rangle_{\rm res}
  =
  \operatorname{Res}_W(fg)
  \label{eq:lg-morse-residue-pairing}
\end{equation}
on \(\operatorname{Jac}(W)\) is nondegenerate.

If \(f=\sum_i h_i\partial_iW\), then \(f(p)=0\) at every critical point
\(p\), so \(\operatorname{Res}_W(f)=0\).  Hence the functional descends to
the quotient.  Nondegeneracy of the Hessian means that each zero of the
gradient map
\[
  \nabla W:\C^n\longrightarrow\C^n,
  \qquad
  \phi\longmapsto(\partial_1W,\ldots,\partial_nW)
\]
is simple.  The local algebra of \(\nabla W\) at such a zero is therefore
\(\C\).  Since the critical set is finite, the Chinese-remainder
decomposition gives
\[
  \operatorname{Jac}(W)\cong
  \prod_{p\in\operatorname{Crit}(W)}\C,
\]
where a class \([f]\) is sent to its values \(f(p)\).  In these coordinates
\eqref{eq:lg-morse-residue-pairing} is the diagonal pairing with nonzero
diagonal entries \(H_W(p)^{-1}\), hence is nondegenerate.

For a compact Riemann surface of genus \(g\), the B-twisted
Landau--Ginzburg zero-mode formula used in this chapter is
\begin{equation}
  \langle f_1\cdots f_m\rangle_g
  =
  \sum_{p\in\operatorname{Crit}(W)}
  \frac{f_1(p)\cdots f_m(p)}
  {H_W(p)^{1-g}}.
  \label{eq:b-twisted-lg-zero-mode-correlator}
\end{equation}
In this monograph \eqref{eq:b-twisted-lg-zero-mode-correlator} has the
following status: it is the finite-dimensional zero-mode and determinant
calculation after the \(Q_B\)-preserving regulator, determinant-line
orientation, and absence of boundary terms have been supplied.  The
continuum QFT assertion is conditional on those analytic inputs.

\section{A-Twisted Sigma-Model Zero Modes}

Let \(X\) be a smooth K\"ahler target with real tangent bundle \(TX\).  The
A-twisted sigma model uses the scalar supercharge
\eqref{eq:a-twist-scalar-supercharge}.  The full model includes nonconstant
maps \(\Sigma\to X\), instanton sectors, compactification of moduli spaces,
and operator contact terms.  The constant-map sector, however, is a
finite-dimensional calculation.

In the constant-map truncation of an A-twisted sigma model with target \(X\),
let \(x^a\) be local real coordinates on \(X\) and let \(\chi^a\) be odd
variables transforming as \(dx^a\).  Suppose the scalar differential acts by
\begin{equation}
  Q_Ax^a=\chi^a,\qquad Q_A\chi^a=0.
  \label{eq:a-model-zero-mode-differential}
\end{equation}
Then the algebra of polynomial functions in the odd variables identifies
with differential forms on \(X\), and \(Q_A\) identifies with the de Rham
differential \(d_X\).  Thus the constant-map local \(Q_A\)-cohomology is
\[
  H^\bullet_{Q_A}\cong H^\bullet_{\rm dR}(X;\C).
\]
Functions of \(x^a\) and the odd variables \(\chi^a\) are local functions on
the parity-shifted tangent bundle \(\Pi TX\).  The coordinate identification
\[
  \chi^{a_1}\cdots\chi^{a_k}
  \longleftrightarrow
  dx^{a_1}\wedge\cdots\wedge dx^{a_k}
\]
identifies these functions with differential forms.  Under this
identification the derivation determined by
\eqref{eq:a-model-zero-mode-differential} is
\[
  Q_A=\chi^a\frac{\partial}{\partial x^a},
\]
which is precisely the de Rham differential in local coordinates.  Its
cohomology is therefore de Rham cohomology after gluing the coordinate
charts.

The nonconstant A-model sectors are classified, at the classical level, by
homology classes \(\beta=\phi_*[\Sigma]\in H_2(X,\mathbb Z)\).  For a
K\"ahler metric with form \(\omega\), the bosonic energy admits the
pointwise Bogomolny decomposition
\begin{equation}
  E(\phi)
  =
  \int_\Sigma |\bar\partial_J\phi|^2\,\mathrm{vol}_\Sigma
  +
  \int_\Sigma\phi^*\omega
  \label{eq:a-model-energy-decomposition}
\end{equation}
in the normalization of the sigma-model action.  Thus holomorphic maps
\(\bar\partial_J\phi=0\) minimize the energy in a fixed homology class.  A
mathematical A-model correlation function requires compactifying the
holomorphic-map moduli space, orienting its virtual determinant line, and
controlling collisions of marked points.  These are theorem-level inputs,
additional to the twist construction.

\section{Abelian Duality And Circle T-Duality}
\label{sec:two-dimensional-abelian-duality}

Abelian duality is used here only as a statement about two-dimensional QFT.
The local calculation is a first-order path integral for a compact scalar
and, in the
\(\mathcal N=(2,2)\) extension, the corresponding Legendre transform between
chiral and twisted-chiral field variables.  The global QFT statement requires
the compact path-integral measure, spin structures, contact terms, and
topological sectors to be specified.

\begin{hypothesis}[Compact scalar first-order duality setup]
\label{hyp:compact-scalar-first-order-duality}
Let \(\Sigma\) be a closed oriented Euclidean surface.  A compact scalar
\(\vartheta:\Sigma\to\mathbb R/2\pi\mathbb Z\) of radius \(R>0\) has action
\begin{equation}
  S_R(\vartheta)
  =
  \frac{R^2}{4\pi}
  \int_\Sigma d\vartheta\wedge *d\vartheta .
  \label{eq:compact-scalar-radius-action}
\end{equation}
The functional integral sums over all winding sectors
\([d\vartheta/2\pi]\in H^1(\Sigma;\mathbb Z)\).  The first-order presentation
uses a real one-form \(A\) with the same integral-period normalization and a
dual periodic scalar \(\widetilde\vartheta\sim
\widetilde\vartheta+2\pi\):
\begin{equation}
  S_{\rm fo}(A,\widetilde\vartheta)
  =
  \frac{R^2}{4\pi}\int_\Sigma A\wedge *A
  +
  \frac{\ii}{2\pi}\int_\Sigma
  \widetilde\vartheta\, dA .
  \label{eq:compact-scalar-first-order-action}
\end{equation}
The measure is assumed to be chosen so that integrating over
\(\widetilde\vartheta\) imposes flatness with integral periods and integrating
over \(A\) is the corresponding Gaussian integral along the standard
Euclidean contour.  Field-independent determinant constants are absorbed in
the normalization of the partition function.
\end{hypothesis}

\paragraph{Compact scalar radius inversion.}
Under Hypothesis~\ref{hyp:compact-scalar-first-order-duality}, the local
first-order duality maps the radius \(R\) compact scalar to a compact scalar
of radius \(R^\vee=R^{-1}\):
\begin{equation}
  S_{R^\vee}(\widetilde\vartheta)
  =
  \frac{1}{4\pi R^2}
  \int_\Sigma d\widetilde\vartheta\wedge *d\widetilde\vartheta .
  \label{eq:compact-scalar-dual-radius-action}
\end{equation}
Applying the same local operation twice returns the original radius.

Integrating over \(\widetilde\vartheta\) in
\eqref{eq:compact-scalar-first-order-action} imposes \(dA=0\) together with
the integral-period condition.  Locally \(A=d\vartheta\), and globally the
integral periods reconstruct the winding sectors of
\eqref{eq:compact-scalar-radius-action}.

For the other elimination, integrate the second term by parts:
\[
  \frac{\ii}{2\pi}\int_\Sigma\widetilde\vartheta\,dA
  =
  -\frac{\ii}{2\pi}\int_\Sigma d\widetilde\vartheta\wedge A .
\]
The Gaussian equation for \(A\) is
\begin{equation}
  R^2 *A=\ii\, d\widetilde\vartheta ,
  \qquad
  A=\frac{\ii}{R^2}*d\widetilde\vartheta ,
  \label{eq:compact-scalar-gaussian-saddle}
\end{equation}
with the usual Euclidean contour understood.  Substituting the saddle and
absorbing the field-independent determinant gives
\eqref{eq:compact-scalar-dual-radius-action}.  Therefore the dual radius is
\(R^{-1}\), and the second duality sends \(R^{-1}\) back to \(R\).

The Hamiltonian spectrum supplies a regulator-independent arithmetic check of
the same radius inversion.  On a spatial circle, write the momentum and
winding labels as
\[
  (n,w)\in\mathbb Z^2 .
\]
In the convention of \eqref{eq:compact-scalar-radius-action}, the left and
right momenta are
\begin{equation}
  p_L(n,w;R)=\frac{n}{R}+wR,\qquad
  p_R(n,w;R)=\frac{n}{R}-wR .
  \label{eq:compact-scalar-left-right-momenta}
\end{equation}
The duality map
\begin{equation}
  R\longmapsto R^{-1},\qquad (n,w)\longmapsto(w,n)
  \label{eq:compact-scalar-momentum-winding-map}
\end{equation}
preserves \(p_L^2+p_R^2\) and flips the sign of \(p_R\):
\begin{equation}
  p_L(w,n;R^{-1})=p_L(n,w;R),\qquad
  p_R(w,n;R^{-1})=-p_R(n,w;R).
  \label{eq:compact-scalar-spectrum-invariance}
\end{equation}
Thus the finite oscillator spectrum and the zero-mode quadratic form are
compatible with the first-order path integral.  A proof of equality of all
correlation functions must realize this exact arithmetic inside a specified
compact path-integral measure and operator-insertion prescription.

\begin{definition}[Local \(\mathcal N=(2,2)\) abelian Legendre duality]
\label{def:n22-abelian-duality-legendre-setup}
Let \(\mathcal B\) be a real \(\mathcal N=(2,2)\) superfield coordinate for a
shift-isometry direction, and let \(Y\) be a twisted-chiral superfield whose
imaginary part is periodic.  Locally, suppressing normalization constants
fixed by the component convention, a first-order superspace functional has
the form
\begin{equation}
  \int d^2x\,d^4\theta\,
  \bigl(\mathcal B\,u-K(\mathcal B)\bigr),
  \qquad u=Y+\bar Y,
  \label{eq:n22-abelian-duality-first-order-kahler}
\end{equation}
where \(K''(\mathcal B)>0\).  Eliminating \(Y\) enforces the local constraint
\(\mathcal B=\Phi+\bar\Phi\) for a chiral field \(\Phi\).  Eliminating
\(\mathcal B\) instead uses
\begin{equation}
  u=K'(\mathcal B)
  \label{eq:n22-abelian-duality-legendre-equation}
\end{equation}
and gives the Legendre-dual K\"ahler potential
\begin{equation}
  \widetilde K(u)=\mathcal B(u)u-K(\mathcal B(u)).
  \label{eq:n22-abelian-duality-legendre-transform}
\end{equation}
\end{definition}

\begin{remark}[Hessian inversion in local abelian duality]
\label{rem:n22-abelian-duality-hessian-inversion}
Assume Definition~\ref{def:n22-abelian-duality-legendre-setup}.  Then
\begin{equation}
  \widetilde K''(u)=\frac{1}{K''(\mathcal B(u))}.
  \label{eq:n22-abelian-duality-hessian-inverse}
\end{equation}
In particular, the free quadratic potential
\[
  K(\mathcal B)=\frac{R^2}{2}\mathcal B^2
\]
is dual to
\[
  \widetilde K(u)=\frac{1}{2R^2}u^2,
\]
the supersymmetric version of the radius inversion
\(R^2\mapsto R^{-2}\).

Differentiate \eqref{eq:n22-abelian-duality-legendre-transform}.  Since
\(u=K'(\mathcal B(u))\),
\[
  \frac{d\widetilde K}{du}
  =
  \mathcal B(u)+u\mathcal B'(u)
  -K'(\mathcal B(u))\mathcal B'(u)
  =
  \mathcal B(u).
\]
Differentiating \eqref{eq:n22-abelian-duality-legendre-equation} gives
\[
  1=K''(\mathcal B(u))\mathcal B'(u),
\]
and hence
\(\widetilde K''(u)=\mathcal B'(u)=1/K''(\mathcal B(u))\).  For
\(K(\mathcal B)=R^2\mathcal B^2/2\), equation
\eqref{eq:n22-abelian-duality-legendre-equation} gives
\(\mathcal B=u/R^2\), and
\eqref{eq:n22-abelian-duality-legendre-transform} gives
\(\widetilde K(u)=u^2/(2R^2)\).
\end{remark}

For a charged chiral field in an abelian GLSM, the same local dualization
couples the dual twisted-chiral variable linearly to the abelian field
strength \(\Sigma\), with coefficient equal to the integer gauge charge.
The perturbative local calculation is therefore compatible with the GLSM charge
matrix.  Nonperturbative exponential terms in the dual twisted
superpotential, when present, require a separate vortex-instanton
calculation: compactness of the vortex moduli space, fermion zero-mode
saturation, determinant-line orientation, and boundary control must be
supplied before such terms are promoted from semiclassical input to a QFT
statement.

\section{Charged-Chiral Dual Variables and Abelian Mirror Presentations}
\label{sec:charged-chiral-dual-mirror-presentations}

The preceding Legendre transform is the local part of the abelian GLSM mirror
construction.  The full mirror statement is a statement about two continuum
quantum field theories, not only about a formal twisted superpotential.  We
therefore name it before using its protected consequences.

\begin{conjecture}[Abelian GLSM/Hori--Vafa mirror QFT equivalence]
\label{conj:abelian-glsm-hori-vafa-mirror-qft}
Fix a two-dimensional \(\mathcal N=(2,2)\) abelian GLSM on a spin worldsheet
with gauge group \(G=U(1)^s/\Gamma\), integral charge matrix \(Q_i{}^a\),
ordinary chiral multiplets \(\Phi_i\), a declared gauge-invariant chiral
superpotential only when its monomials preserve the phase isometries being
dualized or when a separate mirror prescription for the broken isometries is
included, logarithmic FI-theta coordinates \(T_a=2\pi\ii\tau_a\), flavor and
\(R\)-symmetry backgrounds, allowed local counterterms, the compact
gauge-bundle flux lattice, a supersymmetric regulator, and a phase chamber
avoiding the singular FI locus, runaways, and unresolved mixed
Coulomb--Higgs branches.
For \(G=U(1)^s/\Gamma\), the allowed fluxes are the cocharacter lattice
\(\Lambda_G\subset \mathbb R^s\), the theta angles are characters of
\(\Lambda_G\), and the mirror periodicities, allowed exponentials, and any
residual mirror orbifold are part of the same global-form datum.
Let a proposed mirror be a regulated twisted-chiral theory whose field
content includes variables
\[
  Y_i\sim Y_i+2\pi\ii,\qquad \Sigma_a,
\]
global identifications induced by \(G\), vortex-normalization constants
\(c_i\), a mirror \(D\)-term/Kahler functional
\(K_{\rm mir}^{(\Lambda)}(Y,\overline Y,\Sigma,\overline\Sigma)\), a
path-integral measure \({\mathcal D}_{\Lambda}^{\rm mir}\), counterterm and
background-field coordinates, noncompact boundary conditions, and twisted
superpotential of the form
\eqref{eq:glsm-hv-dual-twisted-superpotential}, with the same spin structures
and background fields.  If the data
\[
  \bigl(
  K_{\rm mir}^{(\Lambda)},
  {\mathcal D}_{\Lambda}^{\rm mir},
  \hbox{counterterms},
  \hbox{global/orbifold data},
  \hbox{boundary conditions at infinity}
  \bigr)
\]
are not specified, the displayed Hori--Vafa superpotential is only a
protected twisted-\(F\)-term presentation, not a definition of the mirror QFT.

The conjecture asserts an equivalence of infrared endpoints
\[
  {\rm IR}\bigl({\mathcal T}_{\rm GLSM}^{(\Lambda)}\bigr)
  \simeq
  {\rm IR}\bigl({\mathcal T}_{\rm mir}^{(\Lambda)}
  [K_{\rm mir},{\mathcal D}^{\rm mir},\widetilde W_{\rm dual}]\bigr)
\]
after compatible regulator limits in that chamber.  In this sense their
Hilbert spaces on circles and Hamiltonians agree, their local operator
algebras and OPEs are identified, their A- and B-twisted sectors, pairings,
partition functions, and correlation functions agree including
background-field contact terms, and the equivalence respects line defects,
boundary conditions, topological sectors, theta periods, RG endpoints, and
global gauge data.  Equalities of twisted chiral rings, Coulomb-branch
critical equations, mirror residues, and protected A-twisted correlators are
consequences or evidence for this infrared statement only after their
regulator, operator-map, determinant, \(D\)-term, measure, and
compactification data have been supplied.
\end{conjecture}

\begin{definition}[Full mirror datum versus protected presentation]
\label{def:glsm-mirror-full-qft-datum}
For an abelian GLSM mirror, call
\[
  (Y_i,\Sigma_a,\widetilde W_{\rm dual})
\]
with periodic \(Y_i\) and the charge-linear terms
\(-Q_i{}^a\Sigma_aY_i\) the protected mirror presentation.  A full mirror-QFT
datum is the larger package
\[
  \mathfrak M_\Lambda
  =
  \left(
  K_{\rm mir}^{(\Lambda)},
  {\mathcal D}^{\rm mir}_{\Lambda},
  \widetilde W_{\rm dual}^{(\Lambda)},
  C_{\rm ct}^{(\Lambda)},
  B_{\infty}^{(\Lambda)},
  {\mathcal G}_{\rm glob}^{(\Lambda)},
  R_{\Lambda\to{\rm IR}}
  \right).
\]
Here \(C_{\rm ct}^{(\Lambda)}\) records local counterterm coordinates,
\(B_{\infty}^{(\Lambda)}\) records boundary conditions and allowed integrations
by parts at noncompact or singular ends of field space,
\({\mathcal G}_{\rm glob}^{(\Lambda)}\) records residual orbifolds, spin
structures, theta characters, and background fields, and
\(R_{\Lambda\to{\rm IR}}\) is the Wilsonian scaling map used to compare
operators and states.  Two choices of \(K_{\rm mir}\) with the same
\(\widetilde W_{\rm dual}\) are physically interchangeable only after a
universality or rigidity statement shows that their difference is irrelevant,
redundant, or attracted to the same infrared fixed point with the same
boundary functional \(B_\infty\).  On a noncompact mirror, an integration by
parts that discards a \(D\)-term is valid only when its boundary contribution
at every end has been shown to vanish or has been included as a source/contact
term.
\end{definition}

\begin{definition}[Admissible mirror datum and the IR-equivalence ladder]
\label{def:glsm-admissible-mirror-datum}
The package \(\mathfrak M_\Lambda\) is an admissible mirror datum only after
the symbols in Definition~\ref{def:glsm-mirror-full-qft-datum} have been
realized as cutoff QFT data.  Concretely, for every cutoff scale \(\Lambda\)
one must have a regulated functional integral
\[
  Z_{\Lambda}^{\rm mir}[J]
  =
  \int_{B_\infty^{(\Lambda)}}{\mathcal D}^{\rm mir}_{\Lambda}\Phi\,
  \exp\left(
    -S_{\Lambda,D}^{\rm mir}[\Phi;K_{\rm mir},C_{\rm ct}]
    -S_{\Lambda,F}^{\rm mir}[\Phi;\widetilde W_{\rm dual}]
    +J_A{\mathcal O}_{A,\Lambda}
  \right),
\]
together with a self-adjoint Hamiltonian on the circle, a dense domain fixed by
the noncompact boundary functional \(B_\infty^{(\Lambda)}\), and a renormalized
operator map
\[
  {\mathcal O}_{A}^{\rm IR}
  =
  \lim_{\Lambda\to\infty}
  Z_A(\Lambda)\,{\mathcal O}_{A,\Lambda}
\]
in the topology used to compare correlation functions.  For a noncompact
target this includes the spectral resolution
\[
  {\mathcal H}_{\rm IR}
  =
  \int_{\mathcal C}^{\oplus} d\mu_{\rm cont}(\alpha)\,
  {\mathcal H}_{\alpha}
  \oplus
  \bigoplus_{\beta\in{\mathcal D}}{\mathcal H}_{\beta},
\]
the reflection relation identifying the two asymptotic bases, the residue
normalization of discrete states, and the contact terms carried by
nonnormalizable sources.  These are QFT data, not decorations of
\(\widetilde W\).

There are then three distinct claims.  The existence claim says that an
admissible mirror datum with the protected presentation exists.  The
universality claim says that two admissible data with the same protected
presentation define the same IR theory only when a regulated path between them
keeps the relevant/marginal couplings fixed, has no phase transition or
spectral bifurcation, preserves the reflection phase and pole residues, and
makes every discarded \(D\)-term boundary current vanish or appear as a
matched source term.  The duality claim identifies the IR endpoint of the
original GLSM with that admissible mirror endpoint.  A Hori--Vafa formula
supplies protected coordinates for these claims; it is not itself any one of
the three claims.
\end{definition}

\begin{controlledapproximation}[D-term RG Schur-complement gate]
\label{ca:glsm-mirror-dterm-rg-schur-gate}
The universality part of
Definition~\ref{def:glsm-admissible-mirror-datum} is a Wilsonian statement, not
a slogan about \(D\)-terms.  A finite regulator can see the issue without any
continuum analysis.  Split the mirror Hilbert space into retained and eliminated
subspaces,
\[
  {\mathcal H}_\Lambda=P{\mathcal H}_\Lambda\oplus Q{\mathcal H}_\Lambda,
\]
where \(P\) contains the low states and source rows used in the comparison.
For a candidate mirror datum \(\mathfrak M_\Lambda\), write the finite
Hamiltonian and a source vector in block form
\[
  H_\Lambda=
  \begin{pmatrix}
  H_{PP}&H_{PQ}\\
  H_{QP}&H_{QQ}
  \end{pmatrix},
  \qquad
  o_\Lambda=o_P\oplus o_Q .
\]
At a comparison energy \(E\) below the eliminated spectrum, integrating out the
high block gives the Feshbach--Schur data
\begin{align}
  H_{\rm eff}^{\mathfrak M}(E)
  &=
  H_{PP}+C_{\rm ct}
  -H_{PQ}(H_{QQ}-E)^{-1}H_{QP},
  \label{eq:glsm-mirror-dterm-rg-schur-hamiltonian}\\
  o_{\rm eff}^{\mathfrak M}(E)
  &=
  Z_O\left(
    o_P-H_{PQ}(H_{QQ}-E)^{-1}o_Q
  \right).
  \label{eq:glsm-mirror-dterm-rg-schur-source}
\end{align}
The counterterm \(C_{\rm ct}\) and source renormalization \(Z_O\) are part of
the RG map \(R_{\Lambda\to{\rm IR}}\).  The protected mirror presentation fixes
neither the high-mode coupling \(H_{PQ}\) induced by finite-field Kahler/measure
data nor the eliminated source component \(o_Q\).

A same-endpoint claim for two admissible data \(\mathfrak M_\Lambda\) and
\(\mathfrak M'_\Lambda\) must therefore supply an isometry or source
identification \(U_P\) on the retained space and a bound
\begin{equation}
  \left\|
  H_{\rm eff}^{\mathfrak M}(E)
  -
  U_P^{-1}H_{\rm eff}^{\mathfrak M'}(E)U_P
  \right\|
  +
  \left\|
  o_{\rm eff}^{\mathfrak M}(E)
  -
  U_P^{-1}o_{\rm eff}^{\mathfrak M'}(E)
  \right\|
  \le B_{\rm RG}(E).
  \label{eq:glsm-mirror-dterm-rg-schur-bound}
\end{equation}
For example, in a one-channel finite cell with the same protected low block
\(H_{PP}=2\), the same eliminated energy \(H_{QQ}=5\), and comparison energy
\(E=0\), the high-mode couplings \(H_{PQ}=1\) and \(H_{PQ}=2\) give
\[
  H_{\rm eff}=2-\frac15=\frac95,
  \qquad
  H'_{\rm eff}=2-\frac45=\frac65 .
\]
They are not the same low Hamiltonian until a counterterm \(C_{\rm ct}=3/5\) is
included.  With \(o_P=1\) and \(o_Q=1/2\), the source vectors are
\[
  o_{\rm eff}=\frac9{10},\qquad
  o'_{\rm eff}=\frac45 ,
\]
so even after the Hamiltonian counterterm, the source two-point resolvent
\(|o_{\rm eff}|^2(H_{\rm eff}+1)^{-1}\) is not matched unless the separate
source renormalization \(Z_O=9/8\) is supplied.  Thus finite-field \(D\)-term
and measure choices are harmless only after their Schur complement, counterterm
transport, and source transport have been checked in the same low-energy
observable.
\end{controlledapproximation}

\begin{remark}[Operator sources are not fixed by the spectrum alone]
\label{rem:glsm-mirror-operator-source-data}
The spectral-domain part of
Definition~\ref{def:glsm-admissible-mirror-datum} is necessary, but it does
not by itself identify the local QFT.  In a finite circle regulator, let
\(H_\Lambda e_n=E_n e_n\) and let \({\mathcal O}_{\Lambda}\) be the
renormalized source representative for a local operator.  The source
two-point function contains the matrix-element data
\[
  G_{{\mathcal O},\Lambda}(\beta)
  =
  \langle e_0,
    {\mathcal O}_{\Lambda}^{\dagger}
    e^{-\beta(H_\Lambda-E_0)}
    {\mathcal O}_{\Lambda}
  e_0\rangle
  =
  \sum_n
  |\langle e_n,{\mathcal O}_{\Lambda}e_0\rangle|^2
  e^{-\beta(E_n-E_0)} .
\]
Two candidate mirrors can therefore have the same finite spectrum, the same
protected chiral-ring multiplication, and the same asymptotic \(F\)-term data,
while their source derivatives of \(Z_\Lambda[J]\) differ because the
renormalized operator vector \({\mathcal O}_{\Lambda}e_0\) differs.  A mirror
equivalence must supply the operator topology, source normalization, and
operator/state map that transport these matrix elements.  Protected residues
and chiral-ring identities are valuable tests of that map, not substitutes for
it.
\end{remark}

\begin{controlledapproximation}[Boundary and defect comparison gate]
\label{ca:glsm-mirror-boundary-defect-gate}
The topological and boundary part of a full mirror statement is another
independent QFT probe.  In a finite cylinder regulator, a boundary condition
\(B_\alpha\) is represented in the closed channel by a vector
\(|B_{\alpha,\Lambda}\rangle\) satisfying the regulated gluing equations for the
chosen \(\mathcal N=2\) algebra up to a residual \(R_{\rm gluing}\).  Its
cylinder amplitude is
\begin{equation}
  Z^{\rm cl}_{\alpha\beta,\Lambda}(\ell)
  =
  \langle B_{\alpha,\Lambda}|
  e^{-\ell(H^{\rm cl}_\Lambda-c_\Lambda/12)}
  |B_{\beta,\Lambda}\rangle .
  \label{eq:glsm-boundary-closed-cylinder}
\end{equation}
The same datum must admit an open-channel description
\begin{equation}
  Z^{\rm op}_{\alpha\beta,\Lambda}(t)
  =
  \operatorname{Tr}_{\mathcal H^{\rm op}_{\alpha\beta,\Lambda}}
  e^{-2\pi t(H^{\rm op}_{\alpha\beta,\Lambda}-c_\Lambda/24)}
  \label{eq:glsm-boundary-open-cylinder}
\end{equation}
with the modular relation between \(t\) and \(\ell\), integral open
multiplicities where the regulator has resolved the discrete channels, and a
Cardy residual \(R_{\rm Cardy}\).  A boundary vector component can be orthogonal
to every protected local source row already tested in the closed theory and
still change \eqref{eq:glsm-boundary-closed-cylinder}.  Thus boundary states
are not fixed by the Hamiltonian spectrum, the protected chiral ring, or the
bulk source rows.

Similarly, a topological defect line is represented by an operator
\(D_{r,\Lambda}\) that commutes with the preserved stress tensor and
supercharges up to a residual \(R_{\rm comm}\), has a fusion law
\[
  D_{r,\Lambda}D_{s,\Lambda}
  =
  \sum_u N_{rs}^{u}D_{u,\Lambda}+R^{\rm fus}_{rs,\Lambda},
\]
and has twined traces
\[
  Z_{r,\Lambda}(\beta)
  =
  \operatorname{Tr}_{\mathcal H_\Lambda}
  \bigl(D_{r,\Lambda}e^{-\beta H_\Lambda}\bigr).
\]
Two defects may act identically on the protected chiral subspace while acting
with different phases on non-BPS or continuous states, changing
\(Z_{r,\Lambda}\) and the defect category seen by boundaries.  A regulated
mirror comparison that includes boundaries and defects must therefore carry a
bound of the form
\[
  |\Delta_{\alpha\beta r}|
  \le
  B_{\rm gluing}
  +B_{\rm Cardy}
  +B_{\rm bdry\text{-}src}
  +B_{\rm def\text{-}comm}
  +B_{\rm def\text{-}fus}
  +B_{\rm cont}.
\]
The terms \(B_{\rm bdry\text{-}src}\) and \(B_{\rm def\text{-}fus}\) record,
respectively, boundary-state overlaps invisible to the tested local sources and
defect actions invisible on the protected subspace.  Without this package,
``same superpotential'' and ``same protected local operators'' leave the
boundary/defect sector of the mirror QFT unspecified.
\end{controlledapproximation}

The detailed comparison below is an observable comparison map for
Conjecture~\ref{conj:abelian-glsm-hori-vafa-mirror-qft}.  The physical question
is whether a protected GLSM correlator computed with a regulated vortex measure
agrees with the proposed mirror residue in the same FI, determinant-line, and
operator coordinates.  The object being compared is therefore not the formal
twisted superpotential alone, but the following finite-data package:
\[
  \begin{array}{rcl}
    D_{\rm loc} &:& Y_i\sim Y_i+2\pi\ii,\quad -Q_i{}^a\Sigma_aY_i,\\
    D_{\rm pert} &:& \hbox{Coulomb logarithm and finite FI coordinate},\\
    D_{\rm vort} &:& c_i\hbox{ from determinant, zero modes, orientation, boundary},\\
    D_{\rm obs} &:& \hbox{residue and stable-map comparison.}
  \end{array}
\]
The first two layers are local or protected algebraic checks.  The third layer
is the physical instanton calculation: the vortex collective coordinates are
only the zero-mode part of the problem, while the nonzero-mode determinant,
Berezin saturation, determinant-line orientation, and compactification or
continuum remainder determine whether the coefficient is present and in which
scheme.  The fourth layer compares that coefficient with A-twisted observables.
In this chapter a Hori--Vafa expression is used only after this comparison map
has supplied the direct vortex and observable data it needs.  The remaining
full-QFT proof debts are separate:
\[
  \begin{array}{rcl}
    E_{\rm cont} &:& \hbox{shared continuum construction and Hilbert spaces},\\
    E_{\rm op} &:& \hbox{local operators, OPEs, and source normalizations},\\
    E_{\rm top} &:& \hbox{defects, boundaries, twists, and topological sectors},\\
    E_{\rm bg} &:& \hbox{background fields, contact terms, and theta periods},\\
    E_{\rm sing} &:& \hbox{phase chamber, singular-locus, and RG-endpoint control}.
  \end{array}
\]
A protected-sector equality is recorded with its comparison label rather than
promoted silently to the full conjectural QFT equivalence.

Consider an abelian GLSM with gauge group \(U(1)^s\), twisted chiral field
strengths \(\Sigma_a\), logarithmic FI-theta coordinates \(T_a\), and
chiral multiplets \(\Phi_i\), \(i=1,\ldots,m\), of integer charges
\[
  Q_i{}^a\in\mathbb Z,\qquad a=1,\ldots,s .
\]
For each charged chiral multiplet introduce a periodic twisted chiral
variable
\[
  Y_i\sim Y_i+2\pi\ii .
\]
Repeated abelian gauge indices are summed in this section.  Thus the explicit
formulae below are in the ordinary \(U(1)^s\) covering-torus convention with
flux lattice \(\mathbb Z^s\).  For \(U(1)^s/\Gamma\) they are not obtained by
formal reuse of the same product constraints: one must replace \(\mathbb Z^s\)
by the cocharacter lattice \(\Lambda_G\), make the theta character a character
of \(\Lambda_G\), restrict the allowed dual exponentials to the corresponding
character lattice, and include the residual mirror orbifold.  Until that
lattice datum is supplied, the displayed Hori--Vafa formula is a \(U(1)^s\)
formula.

\begin{controlledapproximation}[Global-form flux lattice gate for abelian
mirrors]
\label{ca:glsm-global-form-flux-lattice-gate}
This warning is not only a matter of notation.  Present the compact gauge
torus as
\[
  G=\mathbb R^s/(2\pi\Lambda_G),
\]
where \(\Lambda_G\) is the cocharacter lattice.  A principal \(G\)-bundle on a
closed oriented surface has magnetic flux
\[
  k=\frac1{2\pi}\int_\Sigma F\in\Lambda_G .
\]
The electric character lattice is the dual lattice
\begin{equation}
  \Lambda_G^\vee
  =
  \{m\in(\mathbb R^s)^*\mid m(k)\in\mathbb Z
    \hbox{ for every }k\in\Lambda_G\}.
  \label{eq:glsm-global-form-character-lattice}
\end{equation}
Every charged chiral field and every allowed dual exponential must carry a
character in \(\Lambda_G^\vee\).  The logarithmic FI coordinate is not just an
element of \(\mathbb C^s\); it is a character of the flux lattice:
\begin{equation}
  \exp\langle T,k\rangle ,
  \qquad
  T\sim T+2\pi\ii\,m,\quad m\in\Lambda_G^\vee .
  \label{eq:glsm-global-form-fi-character}
\end{equation}
These two lattices are what make the theta weight single-valued.  Replacing
\(\Lambda_G\) by \(\mathbb Z^s\) changes the allowed bundles, the theta
periods, and the mirror torus.

For a rank-one illustration, write the covering coordinate so that the
ordinary \(U(1)\) has \(\Lambda=\mathbb Z\).  If the physical gauge group is
the quotient \(U(1)/\mathbb Z_n\), then
\begin{equation}
  \Lambda_G=\frac1n\mathbb Z,
  \qquad
  \Lambda_G^\vee=n\mathbb Z .
  \label{eq:glsm-u1-zn-flux-character-lattices}
\end{equation}
Thus a cover-charge-one chiral multiplet is not a representation of this
quotient group, the minimal flux is \(1/n\), and the logarithmic FI period is
\[
  T\sim T+2\pi\ii n
\]
in the covering coordinate.  A putative \(2\pi\ii\) period would change the
weight of the minimal flux sector by \(\exp(2\pi\ii/n)\).  Conversely, a
charge \(n\) field has integral pairing with every quotient flux, so the local
dual coupling \(-n\Sigma Y\) and the exponential \(e^{-Y}\) are compatible
with the quotient global form.  The quotient may also leave a finite mirror
orbifold, controlled by the finite group
\(\Lambda_G/\mathbb Z^s\) in a chosen covering presentation; its action on the
dual torus is part of the mirror datum, not a convention one can recover from
\(\widetilde W\) after the fact.

Consequently the covering-torus constraints
\(\sum_iQ_i{}^aY_i=-T_a\) and \(q=\exp(T)\) used below are valid only for the
declared \(U(1)^s\) lattice.  For \(U(1)^s/\Gamma\), the same symbolic formula
is meaningful only after \(Q_i\in\Lambda_G^\vee\), \(T\in
{\rm Hom}(\Lambda_G,\mathbb C/2\pi\ii\mathbb Z)\), the allowed exponential
monomials, and the residual mirror orbifold have been supplied.  This is a
physical topological-sector condition, not a harmless rescaling of \(T\).
\end{controlledapproximation}

\begin{controlledapproximation}[Chiral-superpotential phase-isometry gate]
\label{ca:glsm-chiral-superpotential-phase-isometry-gate}
The dual variables \(Y_i\) are introduced by dualizing phase directions of the
charged chiral fields.  A chiral superpotential is therefore admissible for the
displayed abelian mirror formula only along the phase directions that it
actually preserves.  Write
\[
  W_{\rm ch}(\Phi)
  =
  \sum_A \lambda_A\prod_{i=1}^m\Phi_i^{p_{Ai}},
  \qquad p_{Ai}\in\mathbb Z_{\ge0}.
\]
Gauge invariance says
\[
  \sum_i p_{Ai}Q_i{}^a=0
  \qquad\hbox{for every }A,a.
\]
The individual phase rotations used in the local dualization impose the
separate condition below.  If
\[
  \Phi_i\longmapsto e^{\ii v_i\alpha}\Phi_i ,
  \qquad v=(v_i)\in\mathbb Z^m,
\]
then the \(A\)-th monomial acquires the phase
\[
  \exp\!\left(\ii\alpha\sum_i p_{Ai}v_i\right).
\]
Thus the continuous phase lattice available to the first-order duality is
\begin{equation}
  \mathcal L_W
  =
  \left\{
    v\in\mathbb Z^m\;\middle|\;
    \sum_i p_{Ai}v_i=0\ \hbox{for every active monomial }A
  \right\}.
  \label{eq:glsm-superpotential-phase-isometry-lattice}
\end{equation}
Only directions in \(\mathcal L_W\), together with the gauge directions already
accounted for by the \(\Sigma_a\) constraints, can be dualized by the displayed
Legendre-transform argument.  If a monomial violates
\eqref{eq:glsm-superpotential-phase-isometry-lattice}, then one must either
treat \(\lambda_A\) as a background spurion and carry its charge into the mirror
datum, or provide a separate mirror prescription for the broken-isometry
interaction.

For example, with charges \((1,1,-2)\) the monomial \(P X_1X_2\) is
gauge-invariant:
\[
  (1,1,1)\cdot(1,1,-2)=0.
\]
But it is not invariant under the three individual phase rotations of
\((X_1,X_2,P)\).  It preserves only the two-dimensional sublattice
\[
  v_1+v_2+v_P=0 .
\]
The relative phase \(X_1/X_2\) is dualizable, but a mirror formula that simply
adds three unconstrained primitive variables \(Y_{X_1},Y_{X_2},Y_P\) and the
terms \(-Q_i{}^a\Sigma_aY_i-\mu c_i e^{-Y_i}\) has not encoded the chiral
superpotential \(P X_1X_2\).  Hypersurface GLSMs and other models with chiral
superpotentials therefore require their own mirror interaction, orbifold, and
operator-map data; they are not consequences of the bare abelian Hori--Vafa
presentation by charge neutrality alone.
\end{controlledapproximation}

The local first-order functional for one charged chiral variable is the
superspace analogue of the compact scalar construction.  Let \(B_i\) be a
real superfield and set
\begin{equation}
  U_i=Y_i+\overline Y_i .
  \label{eq:glsm-mirror-real-dual-coordinate}
\end{equation}
The \(D\)-term
\begin{equation}
  \int d^2x\,d^4\theta\,
  \left[
    e^{B_i+2Q_i{}^aV_a}
    -
    \frac12 U_iB_i
  \right]
  \label{eq:glsm-charged-chiral-first-order-dterm}
\end{equation}
is invariant under
\[
  V_a\mapsto V_a+\ii(\Lambda_a-\overline\Lambda_a),
  \qquad
  B_i\mapsto
  B_i-2\ii Q_i{}^a(\Lambda_a-\overline\Lambda_a),
  \qquad
  Y_i\mapsto Y_i .
\]
Integrating over \(Y_i\) imposes the twisted linear constraint on \(B_i\).
In components of superspace this is the pair of constraints
\[
  \overline D_+D_-B_i=0,\qquad D_+\overline D_-B_i=0 .
\]
Locally this says
\[
  B_i=\Psi_i+\overline\Psi_i
\]
with \(\Psi_i\) a chiral superfield.  Then
\[
  e^{B_i+2Q_i{}^aV_a}
  =
  \overline\Phi_i
  \exp(2Q_i{}^aV_a)\Phi_i,
  \qquad
  \Phi_i=e^{\Psi_i},
\]
which recovers the original charged chiral kinetic term in that field chart.

Eliminating \(B_i\) instead gives the algebraic equation
\begin{equation}
  e^{B_i+2Q_i{}^aV_a}=\frac12 U_i ,
  \qquad
  B_i=\log\frac{U_i}{2}-2Q_i{}^aV_a .
  \label{eq:glsm-charged-chiral-b-equation}
\end{equation}
Substitution gives a dual \(D\)-term of Legendre form, together with the
linear \(V_aU_i\) term
\[
  Q_i{}^a\int d^2x\,d^4\theta\,V_a(Y_i+\overline Y_i).
\]
Up to affine \(D\)-terms, which change only local counterterm coordinates in
this chart, the classical dual Legendre potential is
\begin{equation}
  K_{i,{\rm dual}}^{\rm cl}(U_i)
  =
  -\frac12 U_i\log\frac{U_i}{2},
  \qquad
  U_i=Y_i+\overline Y_i>0 .
  \label{eq:glsm-charged-chiral-dual-classical-kahler}
\end{equation}
Because the sign of a twisted-chiral \(D\)-term is opposite to the sign of an
ordinary chiral \(D\)-term in the component kinetic convention, the kinetic
coefficient is proportional to
\[
  \frac1{2U_i}.
\]
Thus the dual coordinate lives on a half-cylinder with a singular boundary at
\(U_i=0\), while \(U_i\to+\infty\) is the weak-coupling cylindrical end.  The
classical first-order manipulation therefore supplies not only a holomorphic
coordinate \(X_i=e^{-Y_i}\), but also a domain, a singular boundary, and a
classical measure Jacobian.  In the quantum theory these become part of
\(\mathfrak M_\Lambda\) in
Definition~\ref{def:glsm-mirror-full-qft-datum}: the induced measure must
reproduce the axial anomaly and must specify whether \(D\)-term integrations by
parts have boundary contributions at \(U_i=0\) or at infinity.
The superspace integration-by-parts identity in the component convention of
Remark~\ref{rem:glsm-twisted-f-component-normalization} is
\begin{equation}
  \int d^4\theta\,V_a(Y_i+\overline Y_i)
  =
  -\int d^2\widetilde\theta\,\Sigma_aY_i
  +{\rm c.c.},
  \label{eq:glsm-dual-integration-by-parts}
\end{equation}
where \(Y_i\) is twisted chiral and
\(\Sigma_a=\overline D_+D_-V_a\).  Thus the local dualization produces the
twisted \(F\)-term
\[
  -\int d^2\widetilde\theta\,
  Q_i{}^a\Sigma_aY_i
  +{\rm c.c.}
\]
before any vortex effect is invoked.  This derivation fixes the charge
matrix in the dual variables: the integer \(Q_i{}^a\) that couples
\(\Phi_i\) to \(V_a\) is the same integer multiplying \(\Sigma_aY_i\).
In components, if \(Y_i=y_i+\cdots+\theta^-\overline\theta^+G_i+\cdots\),
the bosonic top component of \(-Q_i{}^a\Sigma_aY_i\) is
\[
  -Q_i{}^a
  \left[
    \sigma_aG_i
    +y_i\bigl(D_a+\ii F_{a,12}\bigr)
  \right],
\]
with the common \(1/(2\pi)\) component pairing of
\eqref{eq:glsm-twisted-f-component-pairing}.  Thus
\(\operatorname{Im}y_i\) is the dynamical theta angle for the same compact
gauge flux that appeared in
Remark~\ref{rem:glsm-compact-fi-theta-normalization}; there is no additional
flavor-labelled flux lattice hiding inside the superspace notation.
The FI term in the same superspace convention is
\[
  -\int d^2\widetilde\theta\,T_a\Sigma_a+{\rm c.c.}
\]
with \(T_a\) the logarithmic coordinate of
Remark~\ref{rem:glsm-compact-fi-theta-normalization}.  Thus the two
periodicities are matched before any Hori--Vafa input is used:
\(T_a\sim T_a+2\pi\ii\), \(Y_i\sim Y_i+2\pi\ii\), and the only integer
multipliers in the local dualization are the gauge charges \(Q_i{}^a\).

The convention in the rest of this section is chosen so that eliminating the
\(Y_i\) matches the one-loop Coulomb superpotential
\eqref{eq:abelian-glsm-one-loop-twisted-superpotential}.  Put
\begin{equation}
  M_i(\Sigma)=\sum_{a=1}^s Q_i{}^a\Sigma_a .
  \label{eq:glsm-mirror-mass-linear-form}
\end{equation}
The dual twisted-superpotential ansatz is
\begin{equation}
  \widetilde W_{\rm dual}(\Sigma,Y)
  =
  -\sum_{a=1}^s T_a\Sigma_a
  -
  \sum_{i=1}^m M_i(\Sigma)Y_i
  -
  \mu\sum_{i=1}^m c_i e^{-Y_i},
  \qquad c_i\in\C^\times .
  \label{eq:glsm-hv-dual-twisted-superpotential}
\end{equation}
The constants \(c_i\) depend on the normalization of the dual variables and
on the vortex determinant convention; rescaling \(e^{-Y_i}\) shifts the
corresponding logarithmic branch and changes \(T_a\) by a finite affine
scheme transformation.  The intrinsic data are the charge matrix, the
periodicity of the \(Y_i\), and the fact that the exponential monomial is
nonzero with the charge dictated by the dualized chiral field.

The local dualization has not derived the exponential term.  It has derived
the periodic variable \(Y_i\), the linear coupling \(-M_iY_i\), and the charge
normalization.  The remaining input is a protected connected amplitude on the
asymptotic end of the dual cylinder.  Because \(Y_i\) is periodic with
imaginary period \(2\pi\ii\), holomorphic local functions on that end are
Laurent series in
\[
  X_i=e^{-Y_i}.
\]
The large positive \(\operatorname{Re}Y_i\) region is the weak-coupling end
of the dualized chiral variable, so regularity there permits nonnegative
powers of \(X_i\).  Suppose, for one dual field and on a simply connected
branch where \(M_i\neq0\), that the connected protected correction is an
ordinary holomorphic germ \(h_i(X_i)\) with \(h_i(0)=0\), so that the relevant
part of the twisted superpotential is
\begin{equation}
  -M_iY_i-h_i(X_i)
  =
  M_i\log X_i-h_i(X_i),
  \qquad X_i=e^{-Y_i}.
  \label{eq:glsm-mirror-general-holomorphic-correction}
\end{equation}
The critical equation is
\begin{equation}
  M_i=X_i h_i'(X_i).
  \label{eq:glsm-mirror-general-critical-equation}
\end{equation}
At a critical point the derivative of the eliminated superpotential is
\begin{equation}
  \frac{\partial \widetilde W_{\rm elim}}{\partial M_i}
  =
  \log X_i(M_i).
  \label{eq:glsm-mirror-eliminated-branch-derivative}
\end{equation}
Indeed, the coefficient of \(\partial X_i/\partial M_i\) is
\(M_i/X_i-h_i'(X_i)\), which vanishes by
\eqref{eq:glsm-mirror-general-critical-equation}.  Exact agreement with the
Coulomb one-loop logarithm
\[
  \frac{\partial}{\partial M_i}
  \left[
    M_i\left(\log\frac{M_i}{\mu c_i}-1\right)
  \right]
  =
  \log\frac{M_i}{\mu c_i}
\]
therefore forces \(X_i(M_i)=M_i/(\mu c_i)\) on the branch.  Substituting into
\eqref{eq:glsm-mirror-general-critical-equation} gives
\(h_i'(X_i)=\mu c_i\), hence
\[
  h_i(X_i)=\mu c_iX_i
\]
up to an irrelevant additive constant.  Thus the primitive monomial
\(-\mu c_i e^{-Y_i}\) is singled out by the combination of periodicity,
regularity at the weak-coupling end, and exact Coulomb-branch matching.  This
argument is not a construction of the amplitude and not a construction of the
mirror \(D\)-term.  It reduces the
nonperturbative theorem boundary to the following sharp questions: construct
the single-vortex sector in the chosen regulator, prove that its fermion zero
modes leave precisely the twisted \(F\)-term measure, prove
\(c_i\neq0\), and justify that possible higher connected harmonics do not
contribute in the protected coordinate being matched.  The separate full-QFT
boundary is to identify the quantum-corrected \(K_{\rm mir}\), induced
measure, allowed \(D\)-term deformations, and noncompact boundary conditions
well enough to show that the protected presentation flows to the claimed
infrared theory.

The next two propositions and the retained-window bound supply the
\(D_{\rm vort}\) part of the observable comparison map for
Conjecture~\ref{conj:abelian-glsm-hori-vafa-mirror-qft}.  They analyze the
coefficient of the primitive mirror monomial as a regulated instanton
amplitude; they supply neither the full continuum operator map nor the
Hilbert-space equivalence by themselves.

\begin{proposition}[Vortex zero-mode filter for twisted \(F\)-terms]
\label{prop:glsm-vortex-zero-mode-filter}
In the finite-regulator charge-one sector used below, let
\(\mathcal D_{F,i,\Lambda}(z)\) be the fermion Hessian at a vortex
configuration \(z\).  Assume its kernel has been split as
\[
  \ker \mathcal D_{F,i,\Lambda}(z)
  =
  \Pi_{\rm univ}(z)\oplus K^{\rm res}_{i,\Lambda}(z),
  \qquad
  \dim \Pi_{\rm univ}=2,
\]
where \(\Pi_{\rm univ}\) is the universal pair identified with the
\(d^2\widetilde\theta\) measure.  Let
\[
  \nu^{\rm res}_{i,\Lambda}(z)=\dim K^{\rm res}_{i,\Lambda}(z).
\]
After the universal pair has been removed, the residual fermion-zero-mode
coefficient has the Berezin form
\begin{equation}
  \mathcal Z^0_{i,F,\Lambda}(z;P)
  =
  \int
  \prod_{\alpha=1}^{\nu^{\rm res}_{i,\Lambda}(z)} \dd\eta_\alpha\,
  P_{\rm ins}(z,\eta),
  \label{eq:glsm-vortex-berezin-zero-mode-filter}
\end{equation}
where \(P_{\rm ins}\) denotes insertions or regulator-allowed zero-mode
liftings.  Hence the charge-one sector can contribute to the uninserted
twisted superpotential only on the locus
\[
  \nu^{\rm res}_{i,\Lambda}(z)=0 .
\]
If \(\nu^{\rm res}_{i,\Lambda}(z)>0\) and no insertion or lifting supplies the
missing residual Grassmann degree, then
\(\mathcal Z^0_{i,F,\Lambda}(z;1)=0\).  The nonzero-mode determinant entering
the coefficient is therefore the determinant on the complement of the full
kernel of \(\mathcal D_{F,i,\Lambda}(z)\), not the raw determinant before
zero-mode removal.
\end{proposition}

\begin{proof}
The two universal zero modes are coordinates on the superspace measure of the
twisted \(F\)-term, so they are not integrated into the scalar coefficient.
All remaining fermion zero modes are ordinary Grassmann variables in the
finite stationary-phase integral.  Berezin integration extracts the coefficient
of the top exterior monomial
\(\eta_1\cdots\eta_{\nu^{\rm res}_{i,\Lambda}(z)}\).  If the insertion
polynomial has smaller residual Grassmann degree, that coefficient is absent
and the integral vanishes.  With no insertion or lifting, the polynomial is
constant, so the only nonzero case is
\(\nu^{\rm res}_{i,\Lambda}(z)=0\).  Removing the full kernel before forming
the determinant is the same collective-coordinate separation on the fermionic
side as the bosonic zero-mode quotient in the vortex measure.
\end{proof}

This is the amplitude question hidden behind the compact mirror formula.  The
integer vortex number is the degree of the common compact gauge bundle.  It is
not a separate topological sector attached to a flavor label.  What may be
indexed by \(i\) is the dual disorder operator \(e^{-Y_i}\), or the source
projection and determinant normalization used to compare the common flux
sector with that primitive mirror monomial.  Coulomb matching fixes the
primitive harmonic once a nonzero coefficient is present.  Neither step proves
that the regulated common vortex sector has exactly the two universal fermion
zero modes or that its determinant line has been oriented consistently.

At finite regulator this distinction is expressed by source projections on one
common sector.  Let \({\mathcal A}_{1,\Lambda}\) denote the
FI-stripped charge-one source functional obtained by integrating over
\(\mathcal U_{1,\Lambda}\), with the gauge orbit, collective coordinates,
ghosts, and universal fermion zero modes already separated.  The
\(i\)-labelled reduced coefficient is a linear projection
\begin{equation}
  \widehat c^{\rm orig}_{i,\Lambda}
  =
  \Pi_{i,\Lambda}\bigl({\mathcal A}_{1,\Lambda}\bigr),
  \label{eq:glsm-common-flux-source-projection}
\end{equation}
where \(\Pi_{i,\Lambda}\) is fixed by the source that will be transported to
the dual disorder operator \(e^{-Y_i}\).  A physical component or A-twisted
observable uses, in general, a different projection
\[
  C_{\mathcal O,\Lambda}^{(1)}
  =
  \Pi_{\mathcal O,\Lambda}\bigl({\mathcal A}_{1,\Lambda}\bigr).
\]
Thus the list of numbers
\(\widehat c^{\rm orig}_{1,\Lambda},\ldots,\widehat c^{\rm orig}_{m,\Lambda}\)
is not itself a construction of a common original-sector amplitude.  In
particular, a product such as \(\prod_i c_{i,\Lambda}^{Q_i{}^a}\) is a
mirror-coordinate assembly after the operator map has been fixed; it is not the
amplitude of one gauge-flux event unless an additional source-factorization or
observable-comparison theorem identifies \(\Pi_{\mathcal O,\Lambda}\) with
that assembled projection and bounds the residuals.  Under a flavor basis
change the projections \(\Pi_i\) transform, while the common sector
\({\mathcal A}_{1,\Lambda}\) does not split into flavor-labelled flux sectors.

\begin{controlledapproximation}[Common-flux operator-map residual diagnostic]
\label{ca:glsm-common-flux-operator-map-diagnostic}
The preceding source-projection formula is still not the operator map.  A
finite-regulator comparison with the primitive mirror operators must supply a
linear source space \(V_{1,\Lambda}\), the common charge-one functional
\({\mathcal A}_{1,\Lambda}\in V_{1,\Lambda}\), original source rows
\(\pi_{i,\Lambda}\), dual source rows \(\widetilde\pi_{i,\Lambda}\), and an
operator transport map
\[
  U_\Lambda:V_{1,\Lambda}\longrightarrow \widetilde V_{1,\Lambda}
\]
such that
\begin{equation}
  \pi_{i,\Lambda}({\mathcal A}_{1,\Lambda})
  =
  \widetilde\pi_{i,\Lambda}
  \bigl(U_\Lambda{\mathcal A}_{1,\Lambda}\bigr)
  +R_{i,\Lambda}^{\rm op}.
  \label{eq:glsm-common-flux-operator-map-row}
\end{equation}
The row \(\widetilde\pi_i\) is the regulated source representative of the
dual monomial \(e^{-Y_i}\).  Equation
\eqref{eq:glsm-common-flux-operator-map-row} is an independent
matrix-element statement; compact flux periodicity and the Coulomb logarithmic
saddle term fix separate parts of the comparison.
For equal-charge fields it must also be covariant under the flavor action:
\[
  U_\Lambda\,\rho_{\rm orig}(g)
  =
  \rho_{\rm mir}(g)\,U_\Lambda,
  \qquad
  \pi_{i,\Lambda}\rho_{\rm orig}(g)
  =
  \sum_j \rho_{ij}^{\vee}(g)\,\pi_{j,\Lambda}.
\]

The same diagnostic has to say which physical observable is being compared.
For an observable row \(\Pi_{\mathcal O,\Lambda}\), the projected coefficients
\(\{\widehat c_{i,\Lambda}^{\rm orig}\}\) determine
\(C_{\mathcal O,\Lambda}^{(1)}\) only if an assembly map
\(\mathfrak A_{\mathcal O,\Lambda}\) is supplied with
\begin{equation}
  \Pi_{\mathcal O,\Lambda}({\mathcal A}_{1,\Lambda})
  =
  \mathfrak A_{\mathcal O,\Lambda}
  \bigl(
    \widehat c_{1,\Lambda}^{\rm orig},\ldots,
    \widehat c_{m,\Lambda}^{\rm orig}
  \bigr)
  +R_{\mathcal O,\Lambda}^{\rm asm}.
  \label{eq:glsm-common-flux-observable-assembly-row}
\end{equation}
If there is a retained direction
\(v\in\bigcap_i\ker\pi_{i,\Lambda}\) but
\(\Pi_{\mathcal O,\Lambda}(v)\ne0\), then two common-flux functionals have the
same primitive mirror coefficients and different observable amplitudes.  In
that case \(\prod_i c_i^{Q_i{}^a}\) is a mirror/FI coordinate, not an
observable amplitude.  A controlled comparison must therefore carry the bound
\[
  |\Delta_{\mathcal O}|
  \le
  B_{\rm op}+B_{\rm src}+B_{\rm span}
  +B_{\rm flav}+B_{\rm asm}+B_{\rm cont},
\]
where \(B_{\rm span}\) controls the failure of the primitive rows to span the
observable row.  The continuum Hori--Vafa operator theorem is precisely the
claim that this package exists, has the declared residuals, and survives the
regulator limit in the chosen source topology.
\end{controlledapproximation}

\begin{proposition}[Finite-regulator vortex fluctuation complex]
\label{prop:glsm-vortex-fluctuation-complex}
At a charge-one vortex saddle \(z\), the coefficient \(c_i\) is controlled by
a gauge-fixed fluctuation complex, not by the vortex collective coordinates
alone.  In a finite supersymmetric regulator, write
\[
  0\longrightarrow
  \mathfrak g_{\Lambda,z}
  \xrightarrow{\ G_z\ }
  \mathcal B_{\Lambda,z}
  \xrightarrow{\ L_z\ }
  \mathcal E_{\Lambda,z}
  \longrightarrow 0,
  \qquad L_zG_z=0,
\]
where \(\mathfrak g_{\Lambda,z}\) is the infinitesimal gauge-orbit space,
\(\mathcal B_{\Lambda,z}\) is the bosonic fluctuation space
\((\delta A,\delta\phi)\), and \(L_z\) is the linearized vortex equation plus
gauge-slice operator.  Let \(\mathcal M_z=G_z^\dagger G_z\) be the
Faddeev--Popov operator on the orbit directions, let \(H_{B,z}\) be the
bosonic Hessian on the orthogonal slice with the vortex collective zero modes
removed, and let \(\mathcal D_{F,z}\) be the fermion Hessian with
\(\Pi_{\rm univ}(z)\oplus K^{\rm res}_{i,\Lambda}(z)\) removed from its
kernel.  Then the nonzero-mode factor entering
\eqref{eq:glsm-single-vortex-original-coefficient-integral} is a determinant-line
element of the form
\begin{equation}
  \mathcal W_{i|1,\Lambda}^{\rm nz}(z)
  =
  \frac{
    \det{}' \mathcal D_{F,z}\,
    \det{}'\mathcal M_z
  }{
    \bigl(\det{}'H_{B,z}\bigr)^{1/2}
  }
  \quad
  \hbox{after pairing with the chosen zero-mode orientations.}
  \label{eq:glsm-vortex-nonzero-mode-determinant-complex}
\end{equation}
A raw determinant over \(\mathcal B_{\Lambda,z}\) or over the full fermion
space vanishes because it includes zero modes.  Omitting
\(\det{}'\mathcal M_z\) changes the gauge-slice density.  Omitting the
orientation of \(\Pi_{\rm univ}\) and \(K^{\rm res}_{i,\Lambda}\) leaves a
line element rather than a scalar coefficient.  Thus the primitive
Hori--Vafa monomial can be used as a protected term only after the
finite-regulator complex, zero-mode removal, determinant-line orientation,
and residual zero-mode saturation have been fixed.
\end{proposition}

\begin{proof}
In finite dimensions the stationary-phase expansion first quotients the gauge
orbit and separates the tangent directions to the vortex moduli.  Gauge
fixing replaces the orbit volume by the Faddeev--Popov determinant
\(\det{}'\mathcal M_z\).  The Gaussian integral over nonzero bosonic normal
directions gives \((\det{}'H_{B,z})^{-1/2}\).  The fermion Gaussian gives
\(\det{}'\mathcal D_{F,z}\) only after its zero modes have been removed from
the determinant and treated as Grassmann collective coordinates.  The two
universal fermion zero modes are identified with the
\(d^2\widetilde\theta\) measure; any residual zero modes are paired with the
Berezin coefficient in
\eqref{eq:glsm-vortex-berezin-zero-mode-filter}.  Multiplying these finite
factors gives
\eqref{eq:glsm-vortex-nonzero-mode-determinant-complex}.  Each prime is
load-bearing: if the bosonic collective modes or fermion zero modes are left
inside the determinant, the determinant is zero rather than a vortex
coefficient.
\end{proof}

\begin{proposition}[Single-vortex coefficient as a protected amplitude]
\label{prop:glsm-single-vortex-exponential-amplitude}
Fix a primitive dual variable \(Y_i\) in the abelian mirror comparison and
choose a supersymmetric finite regulator preserving the compact gauge-bundle
flux and the twisted chiral field strength.  In a rank-one GLSM let
\(\mathcal U_{1,\Lambda}\) be a smooth finite-dimensional chart for the common
charge-one gauge-flux sector, after quotienting by gauge transformations and
after choosing collective-coordinate coordinates \(z\).  For higher rank the
same notation means a chosen primitive flux vector \(k\in\Lambda_G\).  Write
\(\dd\nu_{1,\Lambda}(z)\) for the bosonic zero-mode density,
\(\mathcal W_{i|1,\Lambda}^{\rm nz}(z)\) for the nonzero-mode determinant
ratio in that common vortex background after applying the \(i\)-labelled
source/operator projection, and
\(\mathcal Z_{i,F,\Lambda}^{0}(z)\) for the Berezin coefficient left after
the two universal fermion zero modes have been identified with
\(d^2\widetilde\theta\), as filtered by
Proposition~\ref{prop:glsm-vortex-zero-mode-filter}.  Let
\(q^{\rm top}_{1,\Lambda}\) be the numerical FI-theta weight
\(\exp(T_\Lambda)\) of this common charge-one sector in the original GLSM
regulator.  This is the classical topological sector weight; it is not
flavor-labelled and is not part of the finite determinant or operator
normalization \(c_i\) that later appears in the mirror superpotential.
Finally let
\[
  \mathcal P^{\rm eff}_{i,\Lambda}(z;\xi,\eta,\lambda,J)
  =
  \left\langle
    \exp\!\bigl[-V^{\rm int}_{i,\Lambda,z}(\varphi,\psi)\bigr]
    \mathcal P^{\rm bare}_{\Lambda}
      (z;\xi,\eta,\lambda,J;\varphi,\psi)
  \right\rangle'_{\Lambda,z}
\]
denote the primed normal-mode expectation of the bare zero-mode/source
polynomial together with the cubic and higher interaction vertices after the
Gaussian determinant, zero modes, and ghosts have been separated.  Its
protected zero-mode projection is
\[
  \mathcal Z^{0,{\rm eff}}_{i,F,\Lambda}(z)
  =
  \Bigl[
    \mathcal P^{\rm eff}_{i,\Lambda}(z;\xi,\eta,0,0)
  \Bigr]_{\xi_+\xi_-\,\eta_{\rm top}} .
\]
Only in a localization or protected large-FI window does this effective
projection factor as a source-independent vacuum number
\(\mathcal N^{\rm int}_{i,\Lambda}(z)\) times a Gaussian zero-mode polynomial.
For generic normal-mode interactions, source insertions and interaction
cumulants mix.

After stripping off the classical FI-theta sector weight, the original GLSM
calculation determines the reduced charge-one normalization
\begin{equation}
  \widehat c^{\rm orig}_{i,\Lambda}
  =
  \mu^{-1}
  \int_{\mathcal U_{1,\Lambda}}
  e^{-S^{\rm red}_{1,\Lambda}(z)}\,
  \mathcal W_{i|1,\Lambda}^{\rm nz}(z)\,
  \mathcal Z^{0,{\rm eff}}_{i,F,\Lambda}(z)\,
  \dd\nu_{1,\Lambda}(z).
  \label{eq:glsm-single-vortex-original-coefficient-integral}
\end{equation}
The FI-weighted original-sector contribution is
\(q^{\rm top}_{1,\Lambda}\widehat c^{\rm orig}_{i,\Lambda}\).  Keeping this
product distinct from \(\widehat c^{\rm orig}_{i,\Lambda}\) is essential:
otherwise the same FI-theta weight is counted once in the original vortex
sector and again when the Hori--Vafa constraint produces the exponentiated
mirror coordinate.
The dual Wilsonian statement is obtained only after the abelian dual
operator map sends the common charge-one event, with its \(i\)-labelled
disorder projection, to the twisted-chiral monomial \(e^{-Y_i}\).  In that dual
frame the corresponding term is
\begin{equation}
  \Delta\widetilde W_{i,\Lambda}^{(1)}(Y_i)
  =
  -\mu\,c_{i,\Lambda}\,e^{-Y_i},
  \label{eq:glsm-single-vortex-superpotential-term}
\end{equation}
where \(c_{i,\Lambda}\) is the image of
\(\widehat c^{\rm orig}_{i,\Lambda}\) in the same finite determinant and
operator normalization convention.  Equivalently,
\begin{equation}
  c_{i,\Lambda}
  =
  Z^{\rm map}_{i,\Lambda}\,\widehat c^{\rm orig}_{i,\Lambda},
  \label{eq:glsm-single-vortex-coefficient-integral}
\end{equation}
with \(Z^{\rm map}_{i,\Lambda}=1\) only after the original and dual operator
frames have been normalized identically.  Here
\(S^{\rm red}_{1,\Lambda}\) is the original-regulator action with the
numerical FI-theta sector weight \(q^{\rm top}_{1,\Lambda}\) displayed
separately; \(e^{-Y_i}\) is not inserted into the original path integral.
If additional fermion zero modes remain unsaturated, then
\(\mathcal Z^{0,{\rm eff}}_{i,F,\Lambda}=0\) and this projected common-flux
contribution supplies no twisted-superpotential term.  If the continuum limit of
\eqref{eq:glsm-single-vortex-original-coefficient-integral} exists and is
nonzero, and if the dual operator map has a matching continuum limit, then
the resulting \(c_{i,\Lambda}\) has a limit \(c_i\) in
\eqref{eq:glsm-hv-dual-twisted-superpotential}.
\end{proposition}

\begin{proof}
At fixed regulator the semiclassical expansion around the charge-one vortex
is a finite stationary-phase calculation, but stationary phase is not a
Gaussian identity.  After the collective coordinates, gauge ghosts, and
fermion zero modes are separated, the normal coordinates still carry the
interaction \(V^{\rm int}_{i,\Lambda,z}\).  The Gaussian part supplies
\(\mathcal W^{\rm nz}_{i,1,\Lambda}\), while the remaining normal-mode
expectation acts on the source polynomial.  Two universal fermion zero modes
are precisely the superspace measure of a twisted \(F\)-term; every extra zero
mode must be saturated by an insertion or lifted by an allowed deformation.
Extracting the \(\xi_+\xi_-\eta_{\rm top}\) coefficient gives
\(\mathcal Z^{0,{\rm eff}}_{i,F,\Lambda}\) and hence
\eqref{eq:glsm-single-vortex-original-coefficient-integral}.  The classical
FI-theta factor \(q^{\rm top}_{1,\Lambda}\) multiplies the original common
charge-one sector, but it is stripped before defining the finite
determinant/operator normalization \(c_{i,\Lambda}\).  Abelian dualization is
then a second step: the dual periodic variable records the projected disorder
charge of the common flux event, so the \(i\)-labelled image is \(e^{-Y_i}\),
and the period \(Y_i\sim Y_i+2\pi\ii\) fixes this monomial.  This
produces \eqref{eq:glsm-single-vortex-superpotential-term} only after the
finite operator-map factor in
\eqref{eq:glsm-single-vortex-coefficient-integral} has been fixed.
Multiplying \(e^{-Y_i}\) by a nonzero finite constant rescales the dual
coefficient, which is the same finite normalization freedom recorded above as
an affine shift of the FI-theta coordinate.
\end{proof}

Formula
\eqref{eq:glsm-single-vortex-original-coefficient-integral} isolates the data
hidden in the coefficient \(c_i\).  The original GLSM supplies a numerical
common sector weight, a determinant line, and a source-dependent normal-mode
expectation; the dual frame supplies the operator \(e^{-Y_i}\).  The determinant
ratio, the source-specific interaction cumulants or their localization proof,
the orientation of the two universal fermion zero modes, the saturation of
possible additional zero modes, and the boundary behavior of
\(\mathcal U_{1,\Lambda}\) decide the value and possible vanishing of the
reduced projected coefficient.  Agreement with the Coulomb logarithm is then a protected
consistency test: it fixes the primitive harmonic and exposes any finite
FI-coordinate shift caused by determinant or operator-map normalization,
provided the FI-theta sector weight has not already been folded into \(c_i\).
Subsequent finite-cell formulas keep the older mnemonic subscript \(i,1\) only
as shorthand for this convention: \(1\) is the common flux degree and \(i\) is
the chosen disorder/source projection, not a separate flavor topological
sector.

\begin{controlledapproximation}[One-vortex source functional and
\(F\)-term extraction]
\label{ca:glsm-one-vortex-source-functional-extraction}
The coefficient
\eqref{eq:glsm-single-vortex-original-coefficient-integral} should be read as an
amplitude extraction from a source functional, not as a volume of the vortex
moduli chart.  In the same finite regulator, introduce bookkeeping
Grassmann sources \(\lambda_\pm\) for the two universal fermion zero modes
\(\xi_\pm\), residual Grassmann coordinates \(\eta_\alpha\) for
\(K^{\rm res}_{i,\Lambda}(z)\), and local sources \(J_A\) for the component
operators whose instanton contribution is to be differentiated.  On a
retained common charge-one chart the source polynomial is written in finite
stage form
\begin{equation}
  \mathcal Z^{(1)}_{i,\Lambda}[\lambda,J]
  =
  q^{\rm top}_{1,\Lambda}
  \int_{\mathcal U_{1,\Lambda}}
  e^{-S^{\rm red}_{1,\Lambda}(z)}
  \mathcal W^{\rm nz}_{i|1,\Lambda}(z)
  \dd\nu_{1,\Lambda}(z)
  \int d\xi_+\,d\xi_-\,d\eta\,
  \mathcal P^{\rm eff}_{i,\Lambda}(z;\xi,\eta,\lambda,J).
  \label{eq:glsm-one-vortex-source-functional}
\end{equation}
The effective polynomial contains the zero-mode wavefunctions, the primed
nonzero-mode propagators, and the source-dependent connected interaction
cumulants in the chosen source coordinates.  Replacing it by a vacuum scalar
times a Gaussian source polynomial is an extra theorem, not a definition.
The Wilsonian twisted \(F\)-term projection is the coefficient of
\(\xi_+\xi_-\) and of the top residual \(\eta\)-degree, with the
\(\xi_\pm\) line normalized as the \(d^2\widetilde\theta\) measure:
\begin{equation}
  c_{i,\Lambda}
  =
  Z^{\rm map}_{i,\Lambda}\mu^{-1}
  \int_{\mathcal U_{1,\Lambda}}
  e^{-S^{\rm red}_{1,\Lambda}(z)}
  \mathcal W^{\rm nz}_{i|1,\Lambda}(z)
  \Bigl[
    \mathcal P^{\rm eff}_{i,\Lambda}(z;\xi,\eta,0,0)
  \Bigr]_{\xi_+\xi_-\,\eta_{\rm top}}\,
  \dd\nu_{1,\Lambda}(z).
  \label{eq:glsm-one-vortex-fterm-projection}
\end{equation}
The FI-weighted coefficient in the original source functional is instead
\(q^{\rm top}_{1,\Lambda}\widehat c^{\rm orig}_{i,\Lambda}\).  Equation
\eqref{eq:glsm-one-vortex-fterm-projection} is the reduced coefficient after
the classical sector weight has been assigned to the vortex-number monomial and
the FI constraint.
The ordinary component functional at \(J=0\) is different: if the universal
zero modes are not extracted into \(d^2\widetilde\theta\), the
\(d\xi_+d\xi_-\) integral kills the uninserted component contribution.
Conversely, a two-source component amplitude is obtained by differentiating
\eqref{eq:glsm-one-vortex-source-functional} before setting sources to zero:
\begin{equation}
  \left.
  \frac{\partial^2\mathcal Z^{(1)}_{i,\Lambda}}
       {\partial\lambda_+\partial\lambda_-}
  \right|_{\lambda=J=0}
  =
  q^{\rm top}_{1,\Lambda}
  \int_{\mathcal U_{1,\Lambda}}
  e^{-S^{\rm red}_{1,\Lambda}(z)}
  \mathcal W^{\rm nz}_{i|1,\Lambda}(z)
  \left[
    \frac{\partial^2}
      {\partial\lambda_+\partial\lambda_-}
    \mathcal P^{\rm eff}_{i,\Lambda}(z;\xi,\eta,\lambda,J)
  \right]_{\lambda=J=0;\,\eta_{\rm top}}
  \dd\nu_{1,\Lambda}(z),
  \label{eq:glsm-one-vortex-source-derivative}
\end{equation}
after residual zero modes are saturated.  In the special Gaussian-factorized
case the bracket reduces to
\(\mathcal N^{\rm int}_{i,\Lambda}U_{+,\Lambda}U_{-,\Lambda}
\mathcal Z^{0,{\rm res}}_{i,F,\Lambda}\), but generic higher vertices add
source-dependent connected cumulants.  The functions \(U_{\pm,\Lambda}\) and
those cumulants are part of the physical component amplitude and cannot be
replaced by one unless the source normalization and interaction remainder have
been transported to the same determinant-line convention as
\eqref{eq:glsm-one-vortex-fterm-projection}.

Thus a direct instanton/source comparison has the residual form
\begin{equation}
  A^{(1)}_{i,\Lambda}(\mathcal O)
  -
  q^{\rm top}_{1,\Lambda}
  \widehat c^{\rm orig}_{i,\Lambda}I^{0}_{\mathcal O,\Lambda}
  =
  R_{\rm src}+R_{\rm prop}+R_{\rm det}
  +R_{\rm int}+R_{\rm zm}+R_{\rm cpt}+R_{\rm cont},
  \label{eq:glsm-one-vortex-source-amplitude-telescope}
\end{equation}
where \(I^{0}_{\mathcal O,\Lambda}\) is the retained zero-mode/source
overlap in the normalized convention, \(R_{\rm src}\) records source
wavefunction and operator-coordinate transport, \(R_{\rm prop}\) records
primed nonzero-mode propagator contractions beyond the determinant factor,
\(R_{\rm int}\) records the interacting normal-mode cumulants or their
localization remainder, and the remaining terms have the same determinant, zero-mode,
compactification, and continuum meanings as above.  If the residuals are
bounded by \(B_\alpha\), then
\begin{equation}
  \left|
  A^{(1)}_{i,\Lambda}(\mathcal O)
  -
  q^{\rm top}_{1,\Lambda}
  \widehat c^{\rm orig}_{i,\Lambda}I^{0}_{\mathcal O,\Lambda}
  \right|
  \le
  B_{\rm src}+B_{\rm prop}+B_{\rm det}
  +B_{\rm int}+B_{\rm zm}+B_{\rm cpt}+B_{\rm cont}.
  \label{eq:glsm-one-vortex-source-amplitude-bound}
\end{equation}
After the dual operator map, the corresponding dual-frame comparison replaces
the numerical sector weight by the monomial \(e^{-Y_i}\) and transports
\(\widehat c^{\rm orig}_{i,\Lambda}\) to \(c_{i,\Lambda}\).  Before that map has been
fixed, inserting \(e^{-Y_i}\) into the original source functional is a frame
substitution.  The collective-coordinate chart supplies only the bosonic
integration variables.
The source-differentiated amplitude additionally requires the primed
fluctuation determinant, the ghost density, the zero-mode wavefunctions, the
interacting normal-mode cumulants, the source/operator normalization, and the
residual Berezin gate.  This is why a moduli-space count, a Coulomb logarithm,
or a formal Hori--Vafa exponential cannot by itself replace the regulated
instanton amplitude.
\end{controlledapproximation}

\begin{controlledapproximation}[Finite cell component amplitude from a
one-vortex sector]
\label{ca:glsm-one-vortex-component-amplitude-cell}
The previous source-functional formula becomes a concrete component amplitude
only after the operator sources are projected onto the universal fermion
zero-mode line.  Let a retained finite regulator replace
\(\mathcal U_{1,\Lambda}\) by cells \(p\in P\) with positive weights
\[
  E_p
  =
  e^{-S^{\rm red}_{1,\Lambda}(p)}
  \mathcal W^{\rm nz}_{i|1,\Lambda}(p)
  \nu_p .
\]
For two local component sources \(A,B\), write
\[
  u_A(p)=(u_{A,+}(p),u_{A,-}(p)),\qquad
  u_B(p)=(u_{B,+}(p),u_{B,-}(p))
\]
for their overlaps with the two universal zero modes.  The Berezin extraction
of \(d\xi_+d\xi_-\) gives the oriented minor
\begin{equation}
  \Delta_{AB}(p)
  =
  u_{A,+}(p)u_{B,-}(p)
  -
  u_{A,-}(p)u_{B,+}(p),
  \label{eq:glsm-one-vortex-component-source-minor}
\end{equation}
not the product of norms and not the protected coefficient \(c_i\).  At this
finite regulator the effective two-source insertion is
\begin{equation}
  \mathcal I^{\rm eff}_{AB,\Lambda}(p)
  =
  \left[
    \frac{\partial^2}
      {\partial J_A\partial J_B}
    \mathcal P^{\rm eff}_{i,\Lambda}(p;\xi,\eta,\lambda,J)
  \right]_{\lambda=J=0;\,\xi_+\xi_-\,\eta_{\rm top}} .
  \label{eq:glsm-one-vortex-component-effective-insertion}
\end{equation}
Its Gaussian retained part is
\(\mathcal Z^{0,{\rm res}}_{i,F,\Lambda}(p)
(\Delta_{AB}(p)+C'_{AB}(p))\); generic normal-mode interactions add
source-dependent connected cumulants to
\(\mathcal I^{\rm eff}_{AB,\Lambda}\).  The retained component amplitude is
\begin{equation}
  A^{(1)}_{AB,\Lambda}
  =
  q^{\rm top}_{1,\Lambda}
  \sum_{p\in P}
  E_p\,
  \mathcal I^{\rm eff}_{AB,\Lambda}(p)
  +
  R_{AB,\Lambda},
  \label{eq:glsm-one-vortex-component-cell-amplitude}
\end{equation}
where the contact and interaction pieces are in the same determinant-line
convention as the source coordinates.  Its dual-frame image is obtained only
after the charge-one event and local component sources are transported by the
operator map; in that frame the common monomial is \(e^{-Y_i}\).  The absolute
residual is bounded by
\begin{equation}
  |R_{AB,\Lambda}|
  \le
  B_{\rm src}+B_{\rm prop}+B_{\rm det}+B_{\rm zm}
  +B_{\rm int}+B_{\rm cpt}+B_{\rm op}+B_{\rm cont}.
  \label{eq:glsm-one-vortex-component-cell-bound}
\end{equation}
Several shortcuts are now visibly wrong.  If \(u_A(p)\) and \(u_B(p)\) are
parallel on every retained cell and \(C'_{AB}=0\), the component amplitude
vanishes even when the protected vortex coefficient is nonzero.  If
\(C'_{AB}\) is nonzero, it is a source-specific primed-propagator contribution
and cannot be absorbed into the scalar coefficient \(c_i\).  If the connected
interaction cumulant in
\(\mathcal I^{\rm eff}_{AB,\Lambda}\) is nonzero, a single vacuum factor
\(\mathcal N^{\rm int}\) also fails.  Finally, changing the normalization of
\(u_A,u_B\), the orientation of the zero-mode line, or the operator coordinate
changes \(\Delta_{AB}\); the mirror monomial records the dual image of the
sector but does not supply these original component-source data.
\end{controlledapproximation}

\begin{remark}[Source-frame calibration of a one-vortex
component amplitude]
\label{rem:glsm-one-vortex-source-frame-calibration-map}
The finite component cell also gives a useful calibration test.  Define the
retained protected coefficient and the retained component source integrals by
\[
  C_\Lambda
  =
  \sum_{p\in P}
  E_p
  \Bigl[
    \mathcal P^{\rm eff}_{i,\Lambda}(p;\xi,\eta,0,0)
  \Bigr]_{\xi_+\xi_-\,\eta_{\rm top}},
  \qquad
  J_{AB,\Lambda}
  =
  \sum_{p\in P}E_p\,\mathcal I^{\rm eff}_{AB,\Lambda}(p).
\]
The original retained component amplitude carries the numerical vortex-sector
weight and the source integral:
\[
  A^{(1),{\rm ret}}_{AB,\Lambda}
  =
  q^{\rm top}_{1,\Lambda}J_{AB,\Lambda}
  =
  q^{\rm top}_{1,\Lambda}C_\Lambda\,s_{AB,\Lambda},
  \qquad
  s_{AB,\Lambda}:=\frac{J_{AB,\Lambda}}{C_\Lambda}
\]
when \(C_\Lambda\neq0\).  The number \(s_{AB,\Lambda}\) is a source-frame
quantity: it is made from zero-mode overlaps, primed nonzero-mode contact
terms, and source-dependent interaction cumulants in the same regulator.  It is
not determined by the mirror fugacity.
After the original-to-dual map is fixed, the same common sector factor is
represented in the dual Wilsonian functional by \(e^{-Y_i}\); before that
matching, replacing \(q^{\rm top}_{1,\Lambda}\) by \(e^{-Y_i}\) is a frame
substitution rather than a direct amplitude calculation.

Suppose a reference source pair \(C,D\) has
\(J_{CD,\Lambda}\neq0\).  At the retained level the target/source ratio is
\begin{equation}
  A^{(1),{\rm ret}}_{AB,\Lambda}
  =
  \frac{J_{AB,\Lambda}}{J_{CD,\Lambda}}\,
  A^{(1),{\rm ret}}_{CD,\Lambda}.
  \label{eq:glsm-one-vortex-source-frame-ratio}
\end{equation}
Thus a reference component measurement can calibrate the common vortex
fugacity only after the direct source-frame ratio
\(J_{AB,\Lambda}/J_{CD,\Lambda}\) has been computed.  If
\(J_{CD,\Lambda}\) is small, the same noncancellation discipline as in
Controlled Approximation~\ref{ca:glsm-vortex-coefficient-noncancellation} is
needed for the reference channel.  The retained ratio identity becomes a
physical amplitude estimate only after component and source-frame transport
errors have been estimated in the same regulator convention.  If those
separate estimates are supplied as
\(|R_{AB,\Lambda}|\le B_{AB}\), \(|R_{CD,\Lambda}|\le B_{CD}\), and a
source-frame transport error \(B_{\rm frame}\), the calibrated difference
obeys
\begin{equation}
  \left|
    A^{(1)}_{AB,\Lambda}
    -
    \frac{J_{AB,\Lambda}}{J_{CD,\Lambda}}
    A^{(1)}_{CD,\Lambda}
  \right|
  \le
  B_{AB}
  +
  \left|\frac{J_{AB,\Lambda}}{J_{CD,\Lambda}}\right|B_{CD}
  +
  B_{\rm frame}.
  \label{eq:glsm-one-vortex-source-frame-calibration-bound}
\end{equation}
This equation is the two-dimensional analogue of the four-dimensional
instanton lesson: the topological saddle and its collective-coordinate chart
do not by themselves determine a physical amplitude.  A parallel source pair
has \(J_{AB,\Lambda}=0\) when the primed contact term vanishes, even if
\(C_\Lambda\) and the dual primitive monomial are nonzero.  Conversely, a
nonzero \(C'_{AB}\) changes the component ratio without changing the protected
twisted-superpotential coefficient.  Those data have to be calculated in the
direct vortex fluctuation problem before the mirror expression is used as an
amplitude statement.
\end{remark}

\begin{controlledapproximation}[Single-vortex coefficient noncancellation bound]
\label{ca:glsm-vortex-coefficient-noncancellation}
Fix a compact retained window
\(\mathcal U_R\Subset \mathcal U_{1,\Lambda}\) in the regulated common
charge-one vortex chart and write the reduced scalar integrand in
\eqref{eq:glsm-single-vortex-original-coefficient-integral} as
\[
  \widehat K_{i,\Lambda}(z)
  =
  e^{-S^{\rm red}_{1,\Lambda}(z)}
  \mathcal W_{i|1,\Lambda}^{\rm nz}(z)
  \mathcal Z^{0,{\rm eff}}_{i,F,\Lambda}(z).
\]
Let
\[
  C_R
  =
  \mu^{-1}\int_{\mathcal U_R}\widehat K_{i,\Lambda}(z)\,\dd\nu_{1,\Lambda}(z),
  \qquad
  M_R
  =
  \mu^{-1}\int_{\mathcal U_R}|\widehat K_{i,\Lambda}(z)|\,\dd\nu_{1,\Lambda}(z),
\]
and suppose \(M_R>0\) and
\[
  |C_R|\ge \kappa_R M_R,\qquad 0<\kappa_R\le 1 .
\]
The number \(\kappa_R\) is the retained-window noncancellation margin: it
measures how much of the absolute one-vortex mass survives the determinant
phase, the orientation of the universal fermion-zero-mode line, and the
source-effective residual Berezin coefficient.  Assume further that the
omitted part of the vortex chart, the determinant comparison, the
zero-mode/orientation identification, the compactification boundary, the
operator-map normalization, and the continuum comparison are bounded by
\[
  T_R,\quad
  \epsilon_{\rm det},\quad
  \epsilon_{\rm zm},\quad
  \epsilon_{\rm cpt},\quad
  \epsilon_{\rm map},\quad
  \epsilon_{\rm cont}.
\]
For the reduced original coefficient,
\begin{equation}
  |\widehat c^{\rm orig}_{i,\Lambda}-C_R|
  \le
  T_R+\epsilon_{\rm det}+\epsilon_{\rm zm}
  +\epsilon_{\rm cpt}.
  \label{eq:glsm-vortex-coefficient-noncancellation-bound}
\end{equation}
The FI-weighted original amplitude has the same statement after multiplying
both \(C_R\) and the right-hand side by
\(|q^{\rm top}_{1,\Lambda}|\), plus any separately declared FI-coordinate
transport error.  For the dual-frame coefficient, the comparison target is
\(Z^{\rm map}_{i,\Lambda}C_R\), not \(C_R\):
\begin{equation}
  |c_i^\star-Z^{\rm map}_{i,\Lambda}C_R|
  \le
  |Z^{\rm map}_{i,\Lambda}|
  (T_R+\epsilon_{\rm det}+\epsilon_{\rm zm}+\epsilon_{\rm cpt})
  +\epsilon_{\rm map}
  +\epsilon_{\rm cont},
  \label{eq:glsm-vortex-dual-coefficient-noncancellation-bound}
\end{equation}
where \(c_i^\star=c_{i,\Lambda}\) at fixed regulator with
\(\epsilon_{\rm cont}=0\), and \(c_i^\star=c_i\) when the continuum coefficient
exists and the last comparison error is included.  If
\(Z^{\rm map}_{i,\Lambda}\neq0\) and the right-hand side of
\eqref{eq:glsm-vortex-dual-coefficient-noncancellation-bound} is strictly
smaller than \(|Z^{\rm map}_{i,\Lambda}C_R|\), then \(c_i^\star\neq0\); more
quantitatively,
\[
  \frac{|c_i^\star-Z^{\rm map}_{i,\Lambda}C_R|}
       {|Z^{\rm map}_{i,\Lambda}C_R|}
  \le
  \frac{
    |Z^{\rm map}_{i,\Lambda}|
    (T_R+\epsilon_{\rm det}+\epsilon_{\rm zm}+\epsilon_{\rm cpt})
    +\epsilon_{\rm map}+\epsilon_{\rm cont}
  }{|Z^{\rm map}_{i,\Lambda}|\kappa_R M_R}.
\]
This bound is stronger than charge matching, holomorphy, or the
Coulomb logarithm alone.  Those inputs identify the allowed primitive
monomial and its FI-coordinate normalization.  The bound is the
amplitude-side statement that the regulated determinant line, zero-mode
saturation, source-dependent interaction cumulants, map normalization, and
boundary tail have not made the reduced one-vortex coefficient cancel to zero.
\end{controlledapproximation}

We now move from the \(D_{\rm vort}\) amplitude input to the \(D_{\rm pert}\)
and protected twisted-sector consequences of
Conjecture~\ref{conj:abelian-glsm-hori-vafa-mirror-qft}.  The following
critical-equation and FI-coordinate identities are algebraic consequences of
the mirror presentation once the vortex coefficients and branches have been
supplied.

The \(Y_i\)-critical equation is
\begin{equation}
  \partial_{Y_i}\widetilde W_{\rm dual}=0
  \quad\Longleftrightarrow\quad
  e^{-Y_i}=\frac{M_i(\Sigma)}{\mu c_i}.
  \label{eq:glsm-hv-y-critical-equation}
\end{equation}
On a branch where \(M_i(\Sigma)\neq0\), substituting
\eqref{eq:glsm-hv-y-critical-equation} gives
\begin{equation}
  \widetilde W_{\rm dual}\big|_{\partial_Y\widetilde W=0}
  =
  -\sum_{a=1}^s T_a\Sigma_a
  +
  \sum_{i=1}^m
  M_i(\Sigma)
  \left(
    \log\frac{M_i(\Sigma)}{\mu c_i}-1
  \right).
  \label{eq:glsm-hv-eliminated-one-loop-superpotential}
\end{equation}
For \(s=1\), \(M_i=Q_i\Sigma\), and \(c_i=1\), this is precisely the
charge-resolved form of
\eqref{eq:abelian-glsm-one-loop-twisted-superpotential}, with the constants
\(c_i\) and logarithm branches absorbed into the finite definition of the
FI coordinate.  Thus the charged-chiral dual variables give a nontrivial
consistency check: after the exponential terms have been supplied, their
elimination reproduces the protected Coulomb-branch determinant.

The finite shift of the FI coordinate is explicit.  In the rank-one case,
\eqref{eq:glsm-hv-eliminated-one-loop-superpotential} is
\[
  -T\Sigma
  +
  \sum_iQ_i\Sigma
  \left(
    \log\frac{Q_i\Sigma}{\mu}-1
  \right)
  -
  \Sigma\sum_iQ_i\log c_i .
\]
The last term is affine in \(\Sigma\), hence is the same as replacing
\[
  T\longmapsto T+\sum_iQ_i\log c_i .
\]
This is why the constants \(c_i\) are convention data for the protected
Coulomb matching unless one separately fixes a normalization of the vortex
measure.

\begin{proposition}[Vortex normalization and all-rank FI coordinates]
\label{prop:glsm-vortex-normalization-fi-coordinate}
For an abelian gauge group \(U(1)^s\), fix nonzero vortex coefficients
\(c_i\) and choices of \(\log c_i\).  Set
\[
  \widehat X_i=c_i e^{-Y_i}.
\]
Then the mirror superpotential has normalized primitive terms
\[
  -\mu\sum_i\widehat X_i ,
\]
while the logarithmic-torus constraints are written in the shifted FI
coordinates
\begin{equation}
  \prod_{i=1}^m \widehat X_i^{\,Q_i{}^a}
  =
  \exp\left(T_a+\sum_{i=1}^m Q_i{}^a\log c_i\right),
  \qquad a=1,\ldots,s .
  \label{eq:glsm-all-rank-vortex-normalized-constraints}
\end{equation}
Equivalently, in the Coulomb-eliminated expression the constants \(c_i\)
replace
\begin{equation}
  T_a\longmapsto
  T_a^{\rm phys}
  =
  T_a+\sum_{i=1}^m Q_i{}^a\log c_i .
  \label{eq:glsm-all-rank-vortex-fi-shift}
\end{equation}
These equations define the normalized mirror torus and the affine FI-theta
coordinate used by protected Coulomb/residue calculations.  They do not say
that the original GLSM has \(m\) flavor-labelled degree-one flux events, nor do
they identify \(\exp(T_a)\prod_i c_i^{Q_i{}^a}\) with a direct
source-differentiated amplitude.  That identification belongs to the
observable layer and must pass the projection interface
\eqref{eq:glsm-common-flux-source-projection}.
\end{proposition}

\begin{proof}
The constraint \(\sum_i Q_i{}^aY_i=-T_a\) is equivalent, on the logarithmic
torus \(X_i=e^{-Y_i}\), to
\[
  \prod_iX_i^{Q_i{}^a}=\ee^{T_a}.
\]
Multiplying \(X_i\) by \(c_i\) gives
\[
  \prod_i\widehat X_i^{Q_i{}^a}
  =
  \left(\prod_iX_i^{Q_i{}^a}\right)
  \left(\prod_i c_i^{Q_i{}^a}\right),
\]
which is \eqref{eq:glsm-all-rank-vortex-normalized-constraints}.  The same
shift appears before imposing the low-energy constraint.  Since
\[
  \sum_i M_i(\Sigma)\log c_i
  =
  \sum_i\sum_a Q_i{}^a\Sigma_a\log c_i
  =
  \sum_a\Sigma_a\sum_iQ_i{}^a\log c_i ,
\]
the \(c_i\)-dependent part of
\eqref{eq:glsm-hv-eliminated-one-loop-superpotential} is
\(-\sum_a\Sigma_a\sum_iQ_i{}^a\log c_i\), precisely the affine FI-theta
shift in \eqref{eq:glsm-all-rank-vortex-fi-shift}.
\end{proof}

Thus the Hori--Vafa expression is not a normalization-free statement of a
bare path integral.  Its invariant protected input is the charge lattice, the
primitive nonzero vortex monomial for each allowed charge-one sector, and the
FI-theta coordinate after the finite determinant normalization has been
chosen.  Changing the regulator normalization rescales the mirror torus
coordinates and moves the same finite data into \(T_a^{\rm phys}\); it does
not by itself prove the existence or nonvanishing of the vortex coefficient.
Changing the branch by \(\log c_i\mapsto\log c_i+2\pi\ii n_i\) changes
\(T_a^{\rm phys}\) by
\[
  2\pi\ii\sum_iQ_i{}^a n_i ,
  \qquad n_i\in\mathbb Z ,
\]
which is an integral theta-period shift.  The exponentiated constraints
\eqref{eq:glsm-all-rank-vortex-normalized-constraints} are therefore the
branch-independent protected statement.

\begin{controlledapproximation}[Vortex fugacity and dimensional transmutation]
\label{ca:glsm-vortex-fugacity-transmutation}
For a rank-one abelian GLSM with nonzero charges \(Q_i\), put
\[
  Q_\Sigma=\sum_i Q_i,\qquad
  C_Q^{\rm br}
  =
  \exp\!\left(\sum_i Q_i\log_{\rm br}Q_i\right),
\]
where the logarithm branches are the same branches used in the Coulomb
determinant.  The vortex-normalized exponentiated FI coordinate at regulator
scale \(\mu\) is
\begin{equation}
  q_{\rm phys}(\mu)
  =
  \exp\!\left(T(\mu)+\sum_iQ_i\log c_i(\mu)\right).
  \label{eq:glsm-vortex-normalized-fugacity}
\end{equation}
After the \(Y_i\) variables are eliminated, the Coulomb critical equation is
therefore
\begin{equation}
  C_Q^{\rm br}\,\mu^{-Q_\Sigma}\sigma^{Q_\Sigma}
  =
  q_{\rm phys}(\mu).
  \label{eq:glsm-vortex-normalized-coulomb-scale-equation}
\end{equation}
If \(Q_\Sigma\neq0\), the protected scale coordinate is the
renormalization-group invariant combination
\begin{equation}
  \Lambda_Q^{\,Q_\Sigma}
  =
  \frac{\mu^{Q_\Sigma}q_{\rm phys}(\mu)}{C_Q^{\rm br}},
  \qquad
  \sigma^{Q_\Sigma}=\Lambda_Q^{\,Q_\Sigma}.
  \label{eq:glsm-vortex-dimensional-transmutation-scale}
\end{equation}
Changing the regulator scale or the finite vortex determinant normalization
does not preserve the protected Coulomb roots unless
\begin{equation}
  \frac{d}{d\log\mu}
  \left(
    T(\mu)+\sum_iQ_i\log c_i(\mu)+Q_\Sigma\log\mu
  \right)
  =
  0
  \label{eq:glsm-vortex-fugacity-rg-invariance}
\end{equation}
on the selected branch.  If \(Q_\Sigma=0\), the \(\mu\)-power cancels instead:
the same finite determinant constants only define a dimensionless
exponentiated FI coordinate.

This is the field-theoretic content behind the statement that the scale
\(\mu\) in the mirror superpotential is fake.  It is fake only after the
one-loop Coulomb determinant, the finite vortex coefficient, and the FI
counterterm have been combined into
\eqref{eq:glsm-vortex-normalized-fugacity}--\eqref{eq:glsm-vortex-dimensional-transmutation-scale}.
Using the bare \(\ee^T\) while dropping the \(c_i\), or varying \(\mu\) without
the compensating FI flow, changes the Coulomb roots and hence changes the
protected massive vacua.  Conversely, even an invariant
\(\Lambda_Q^{Q_\Sigma}\) proves only the protected twisted-\(F\)-term
coordinate; it does not prove by itself that the full GLSM and the
Landau--Ginzburg mirror are equivalent as continuum QFTs.
\end{controlledapproximation}

There is a complementary low-energy use of the same expression.  In a
regime where the abelian vector multiplet has no independent low-energy
propagating mode, \(\Sigma_a\) acts as a twisted chiral Lagrange multiplier.
The \(\Sigma_a\)-critical equations of
\eqref{eq:glsm-hv-dual-twisted-superpotential} are
\begin{equation}
  \sum_{i=1}^m Q_i{}^aY_i=-T_a,
  \qquad a=1,\ldots,s .
  \label{eq:glsm-hv-sigma-constraint}
\end{equation}
The dual presentation is then a Landau--Ginzburg model on the logarithmic
torus with coordinates \(Y_i\), subject to the linear constraints
\eqref{eq:glsm-hv-sigma-constraint}, and with twisted superpotential
\[
  -\mu\sum_i c_i e^{-Y_i}
\]
restricted to that subspace.  This is a presentation of protected
twisted-chiral data.  A full mirror equivalence requires a regulator,
identification of the local and extended operator algebras, control of the
singular loci and topological sectors, and convergence of the two continuum
limits.

The projective-space calculations below supply the \(D_{\rm obs}\) protected
observable comparison.  They provide evidence for
Conjecture~\ref{conj:abelian-glsm-hori-vafa-mirror-qft} by testing the
vortex-normalized mirror against A-twisted residue and stable-map data after
the operator coordinate and degree-one instanton coefficient have been fixed.

\begin{example}[\(\mathbb P^{N-1}\) dual critical points]
\label{ex:cpn-hv-dual-critical-points}
Take one \(U(1)\) gauge factor and \(N\) chiral multiplets of charge \(1\).
The constraint \eqref{eq:glsm-hv-sigma-constraint} is
\[
  Y_1+\cdots+Y_N=-T .
\]
Set \(X_i=e^{-Y_i}\).  Then
\[
  X_1\cdots X_N=\ee^T .
\]
For \(c_i=1\), the constrained twisted superpotential is proportional to
\[
  X_1+\cdots+X_N
\]
on the algebraic torus with fixed product \(\ee^T\).  Eliminating
\[
  X_N=\frac{\ee^T}{X_1\cdots X_{N-1}}
\]
gives the critical equations
\[
  X_a=\frac{\ee^T}{X_1\cdots X_{N-1}},
  \qquad a=1,\ldots,N-1.
\]
Thus all \(X_i\) are equal and
\[
  X_i^N=\ee^T.
\]
For \(\ee^T\neq0\) there are \(N\) simple critical points, in agreement with
the Coulomb-branch vacuum count in
Example~\ref{ex:cpn-glsm-coulomb-vacua}.  The agreement is algebraic; it is
the protected finite calculation that a full path-integral equivalence must
reproduce after its regulator, local operators, and topological sectors have
been specified.
\end{example}

\begin{proposition}[\(\mathbb P^{N-1}\) mirror residue and the quantum-product test]
\label{prop:cpn-hv-residue-quantum-product}
For the charge-one \(\mathbb P^{N-1}\) GLSM, use the normalized mirror
variables of Proposition~\ref{prop:glsm-vortex-normalization-fi-coordinate}
and write
\begin{equation}
  q_{\rm phys}
  =
  \exp\left(T+\sum_{i=1}^N\log c_i\right)
  \neq0 .
  \label{eq:cpn-hv-physical-q}
\end{equation}
Here \(q_{\rm phys}\) is the mirror/FI coordinate after determinant and
operator normalizations have been transported into the logarithmic torus.  It
is not, by definition, the direct amplitude of a single common original flux
sector.  The latter must be computed with the source projections and residual
budgets above, and only then compared with the quantum-product observable.
On the logarithmic torus
\[
  \prod_{i=1}^N\widehat X_i=q_{\rm phys},
  \qquad
  \widetilde W=-\mu\sum_{i=1}^N\widehat X_i ,
\]
let \(H\) denote the protected hyperplane/Coulomb class represented in the
mirror Jacobi quotient by \([\widehat X_1]\).  With the determinant-line
normalization that strips the universal factor \((-\mu)^{N-1}\) from the
logarithmic Hessian, the mirror residue trace is
\begin{equation}
  \mathcal R_{\mathbb P^{N-1}}(H^k)
  =
  \sum_{x^N=q_{\rm phys}}
  \frac{x^k}{N x^{N-1}}
  =
  \begin{cases}
    q_{\rm phys}^{\,d}, & k=N-1+dN,\quad d\ge0,\\
    0, & k\not\equiv N-1\pmod{N} .
  \end{cases}
  \label{eq:cpn-hv-residue-selection-rule}
\end{equation}
Consequently the protected Frobenius algebra seen by this trace has
\[
  H^N=q_{\rm phys},
  \qquad
  \mathcal R_{\mathbb P^{N-1}}(H^{N-1})=1 .
\]
\end{proposition}

\begin{proof}
Use logarithmic coordinates
\[
  u_a=\log\widehat X_a,\qquad a=1,\ldots,N-1,
\]
and eliminate
\[
  \widehat X_N=q_{\rm phys}\exp\left(-\sum_{a=1}^{N-1}u_a\right).
\]
For the reduced function
\[
  S(u)=\sum_{a=1}^{N-1}e^{u_a}
  +
  q_{\rm phys}\exp\left(-\sum_{a=1}^{N-1}u_a\right),
\]
the critical equations are
\[
  e^{u_a}=\widehat X_N,\qquad a=1,\ldots,N-1.
\]
Thus every critical point has
\[
  \widehat X_1=\cdots=\widehat X_N=x,
  \qquad
  x^N=q_{\rm phys}.
\]
At such a point the logarithmic Hessian of \(S\) is
\[
  x(I_{N-1}+J_{N-1}),
\]
where \(J_{N-1}\) is the all-ones matrix.  Since \(I_{N-1}+J_{N-1}\) has
eigenvalues \(1,\ldots,1,N\), its determinant is \(N x^{N-1}\).  The factor
from \(\widetilde W=-\mu S\) is the universal \((-\mu)^{N-1}\) determinant-line
factor already fixed in the trace normalization, leaving the denominator in
\eqref{eq:cpn-hv-residue-selection-rule}.

It remains to evaluate the finite root sum.  Put
\[
  \ell=k-(N-1).
\]
Then
\[
  \sum_{x^N=q_{\rm phys}}\frac{x^k}{N x^{N-1}}
  =
  \frac1N\sum_{x^N=q_{\rm phys}}x^\ell .
\]
If \(\ell\) is not divisible by \(N\), the sum over the \(N\)-th roots vanishes.
If \(\ell=dN\) with \(d\ge0\), each critical point contributes \(x^{dN}
=q_{\rm phys}^{\,d}\), and the prefactor \(1/N\) cancels the number of roots.
The same critical-point equations identify \(H^N\) and \(q_{\rm phys}\) in the
Jacobi quotient; the nondegenerate residue pairing separates the quotient
classes.
\end{proof}

The status of Proposition~\ref{prop:cpn-hv-residue-quantum-product} is a
protected observable test of the vortex-normalized mirror, not a standalone
proof of Conjecture~\ref{conj:abelian-glsm-hori-vafa-mirror-qft}.  Once a
compactified A-twisted GLSM or
sigma-model construction identifies \(H\) with the hyperplane class and
\(q_{\rm phys}\) with the degree-one vortex amplitude, the formula becomes the
primary selection rule
\[
  \langle H^{N-1+dN}\rangle_{0}=q_{\rm phys}^{\,d},
  \qquad
  \langle H^k\rangle_0=0
  \quad(k\not\equiv N-1\!\!\pmod{N}).
\]
The mirror residue supplies the finite Hessian and root-sum side of that
statement.  The physical identification still depends on the same determinant,
orientation, zero-mode, compactification, and contact-term data isolated in
\eqref{eq:glsm-single-vortex-original-coefficient-integral}.

The following finite map is the same \(D_{\rm obs}\) comparison written at
finite regulator.  Its residual terms are proof obligations, not evidence by
themselves: they must be constructed and controlled before the mirror residue
is a protected A-twisted QFT correlator.

\begin{remark}[From vortex coefficient to protected observable]
\label{rem:glsm-vortex-to-observable-proof-obligation-map}
In the charge-one \(\mathbb P^{N-1}\) GLSM, fix a finite supersymmetric
regulator and let \(\widehat c^{\rm orig}_{i,\Lambda}\) be the reduced
original-GLSM coefficients in
\eqref{eq:glsm-single-vortex-original-coefficient-integral}, with the
classical FI-theta sector weights divided out.  After the original-to-dual
operator map, write
\(c_{i,\Lambda}=Z^{\rm map}_{i,\Lambda}\widehat c^{\rm orig}_{i,\Lambda}\).
Put
\[
  q_\Lambda
  =
  \exp(T)\prod_{i=1}^N c_{i,\Lambda}.
\]
For \(k\ge0\), define the finite mirror-residue prediction
\[
  S_k(q_\Lambda)
  =
  \begin{cases}
    q_\Lambda^{\,d}, & k=N-1+dN,\quad d\ge0,\\
    0, & k\not\equiv N-1\pmod N .
  \end{cases}
\]
Let \(A_{\Lambda,k}^{A}\) be the corresponding regulated A-twisted
correlator after the local operator assigned to \(H^k\) has been chosen.
Suppose the comparison is organized through the stage residuals
\begin{equation}
  A_{\Lambda,k}^{A}-S_k(q_\Lambda)
  =
  R_{\rm coeff}
  +R_{\rm det}
  +R_{\rm zm}
  +R_{\rm cpt}
  +R_{\rm gl}
  +R_{\rm op}
  +R_{\rm cont}.
  \label{eq:glsm-vortex-to-observable-residual-telescope}
\end{equation}
Here \(R_{\rm coeff}\) measures the difference between the continuum
charge-one vortex amplitude and the regulated coefficient
\eqref{eq:glsm-single-vortex-original-coefficient-integral},
\(R_{\rm det}\) the nonzero-mode determinant and determinant-line
normalization error, \(R_{\rm zm}\) the residual zero-mode saturation and
source-insertion error, \(R_{\rm cpt}\) the compactification, obstruction,
boundary, and contact-term error, \(R_{\rm gl}\) the multi-vortex or
degree-\(d\) gluing error when \(d>1\), \(R_{\rm op}\) the map from the GLSM
operator to the hyperplane/Coulomb class, and \(R_{\rm cont}\) the continuum
limit remainder.  These names do not by themselves supply estimates: the
degree-one incidence calculation controls only the finite retained
coefficient, while the determinant, zero-mode, compactification, gluing,
operator-map, and continuum terms remain imported or model-specific
obligations unless constructed separately.  If
\[
  |R_\alpha|\le \epsilon_\alpha
  \quad
  (\alpha={\rm coeff,det,zm,cpt,gl,op,cont}),
\]
then
\begin{equation}
  \left|A_{\Lambda,k}^{A}-S_k(q_\Lambda)\right|
  \le
  \epsilon_{\rm coeff}
  +\epsilon_{\rm det}
  +\epsilon_{\rm zm}
  +\epsilon_{\rm cpt}
  +\epsilon_{\rm gl}
  +\epsilon_{\rm op}
  +\epsilon_{\rm cont}.
  \label{eq:glsm-vortex-to-observable-residual-bound}
\end{equation}
Consequently the quantum-product statement \(H^N=q_{\rm phys}\) follows from
the mirror residue only after \(q_\Lambda\to q_{\rm phys}\neq0\) and all
seven residuals vanish in the protected comparison.  If some
\(c_{i,\Lambda}=0\) because the determinant line is unoriented, an extra
fermion zero mode is unsaturated, or the compactified vortex integral
vanishes, then \(q_\Lambda=0\); the nonzero degree-one quantum product cannot
be recovered by holomorphy or charge matching alone.
\end{remark}

\paragraph{Status of the telescope.}
The identity
\eqref{eq:glsm-vortex-to-observable-residual-telescope} is obtained by
adding and subtracting the successive finite approximants: the regulated
vortex coefficient, the determinant-normalized coefficient, the
zero-mode-saturated coefficient, the compactified A-model amplitude, the
glued degree-\(d\) sector, the local-operator identification, and the
continuum protected correlator.  The estimate
\eqref{eq:glsm-vortex-to-observable-residual-bound} is the triangle
inequality.  Proposition~\ref{prop:cpn-hv-residue-quantum-product} supplies
the finite residue value \(S_k(q_\Lambda)\), but it contains no information
about whether the regulated vortex coefficient is nonzero or whether the
comparison residuals vanish.  That information is precisely the physical
amplitude input isolated above.

The next two statements give the direct stable-map side of the same protected
observable comparison for
Conjecture~\ref{conj:abelian-glsm-hori-vafa-mirror-qft}.  They compute the
incidence and degree constraints that the GLSM vortex comparison must match,
while keeping determinant, zero-mode, compactification, and local-operator
input separate.

\begin{proposition}[Stable-map computation for the \(\mathbb P^{N-1}\) quantum product]
\label{prop:cpn-stable-map-quantum-product-gate}
Let \(N\ge2\).  Work in the A-twisted positive chamber of the
\(\mathbb P^{N-1}\) GLSM, or equivalently in the large-volume sigma-model
chart, and assume the compactified genus-zero stable-map moduli spaces
\[
  \overline{\mathcal M}_{0,3}(\mathbb P^{N-1},d)
\]
come with their virtual fundamental classes, determinant-line orientation,
evaluation maps \({\rm ev}_j\), and the operator map identifying \(H\) with
the hyperplane class.  Then
\[
  {\rm vdim}_{\mathbb C}\,
  \overline{\mathcal M}_{0,3}(\mathbb P^{N-1},d)
  =
  (N-1)+Nd .
\]
For \(d=1\), the three-point invariant
\begin{equation}
  I_1(H,H^{N-1},H^{N-1})
  =
  \int_{[\overline{\mathcal M}_{0,3}(\mathbb P^{N-1},1)]^{\rm vir}}
  {\rm ev}_1^\ast H\,
  {\rm ev}_2^\ast H^{N-1}\,
  {\rm ev}_3^\ast H^{N-1}
  \label{eq:cpn-degree-one-stable-map-line-count}
\end{equation}
equals \(1\), with the complex orientation.  Consequently the degree-one
part of the A-model product gives
\[
  H\star H^{N-1}=q_{\rm phys}\,1,
  \qquad
  H^N=q_{\rm phys}
\]
provided the degree-one stable-map/vortex coefficient is the same
\(q_{\rm phys}\) used in the normalized mirror coordinate.  If the virtual
class, determinant orientation, compactification boundary terms, or operator
map are not fixed, this implication is only a target comparison, not a
derivation from the mirror residue.
\end{proposition}

\begin{proof}
The expected-dimension formula for genus-zero stable maps gives
\[
  {\rm vdim}_{\mathbb C}
  =
  \dim_{\mathbb C}\mathbb P^{N-1}
  +c_1(T\mathbb P^{N-1})\cdot d
  =
  (N-1)+Nd
\]
for three marked points.  At \(d=1\), the insertion in
\eqref{eq:cpn-degree-one-stable-map-line-count} has complex codimension
\[
  1+(N-1)+(N-1)=2N-1,
\]
which is exactly the virtual dimension.  Geometrically, the two
\(H^{N-1}\) insertions constrain the image line to pass through two generic
points of \(\mathbb P^{N-1}\).  There is a unique such line.  A generic
hyperplane is not that line and intersects it once, giving the first marked
point uniquely.  The complex orientation gives positive sign, so the count is
one.

For the quantum product,
\[
  \eta(\alpha\star\beta,\gamma)
  =
  \sum_{d\ge0}q_{\rm phys}^{\,d} I_d(\alpha,\beta,\gamma).
  \]
Taking \(\alpha=H\), \(\beta=H^{N-1}\), and
\(\gamma=H^a\), the dimension constraint is
\[
  1+(N-1)+a=N-1+Nd .
\]
With \(0\le a\le N-1\), it is satisfied only for \(d=1\) and
\(a=N-1\).  Thus \(H\star H^{N-1}\) pairs with the top class exactly as
\(q_{\rm phys}1\) does, and it pairs trivially with the other basis classes.
Nondegeneracy of the classical pairing gives the product relation.
\end{proof}

\begin{controlledapproximation}[Degree-one A-model zero-mode measure bridge]
\label{ca:cpn-degree-one-amodel-zero-mode-measure-bridge}
The line count in Proposition~\ref{prop:cpn-stable-map-quantum-product-gate}
is the geometric part of the A-twisted path integral, not the whole
instanton calculation.  In a finite supersymmetric regulator for the
positive-chamber \(\mathbb P^{N-1}\) GLSM, write the degree-one contribution
to a three-point A-twisted correlator in the stage form
\begin{equation}
  \mathcal C_{\Lambda,1}^{A}(\mathcal O_1,\mathcal O_2,\mathcal O_3)
  =
  q_\Lambda
  \int_{\mathcal U_{\Lambda,1}}
    \mathcal W_{\Lambda,1}^{\rm nz}(z)\,
    \mathcal Z_{\Lambda,1}^{0}(z;\mathcal O_1,\mathcal O_2,\mathcal O_3)\,
    \dd\nu_{\Lambda,1}(z)
  +R_{\rm cpt}+R_{\rm cont}.
  \label{eq:cpn-degree-one-amodel-zero-mode-integral}
\end{equation}
Here \(q_\Lambda=\exp(T)\prod_i c_{i,\Lambda}\) is the supplied
vortex-normalized fugacity; the \(c_{i,\Lambda}\) are reduced
determinant/operator normalizations and do not include the classical FI-theta
sector weight already represented by \(\exp(T)\).  The density
\(\dd\nu_{\Lambda,1}\) is the bosonic collective-coordinate density after gauge
quotient, \(\mathcal W_{\Lambda,1}^{\rm nz}\) is the nonzero-mode
determinant-line element after the tangent zero modes have been removed, and
\(\mathcal Z_{\Lambda,1}^{0}\) is the Berezin coefficient from the A-twisted
fermion zero modes and the three operator insertions.

For a convex target such as \(\mathbb P^{N-1}\), the degree-one obstruction
space \(H^1(\mathbb P^1,\phi^\ast T\mathbb P^{N-1})\) vanishes.  Thus, after
the virtual orientation and the finite GLSM determinant-line comparison have
been fixed, the fermion zero modes identify with the tangent directions of
\(\overline{\mathcal M}_{0,3}(\mathbb P^{N-1},1)\).  The Berezin coefficient
then extracts the top exterior degree:
\begin{equation}
  \mathcal Z_{\Lambda,1}^{0}
  \quad\leadsto\quad
  {\rm ev}_1^\ast\alpha_1\,
  {\rm ev}_2^\ast\alpha_2\,
  {\rm ev}_3^\ast\alpha_3
  \quad
  \hbox{only if }
  \sum_j\operatorname{codim}_{\mathbb C}\alpha_j=2N-1 .
  \label{eq:cpn-degree-one-amodel-berezin-degree}
\end{equation}
For
\((\alpha_1,\alpha_2,\alpha_3)=(H,H^{N-1},H^{N-1})\), this degree is
\(2N-1\), the same as the affine incidence chart in
Remark~\ref{rem:cpn-degree-one-vortex-observable-proof-obligation-template}.
If the finite determinant-line comparison gives orientation sign \(+1\) and
no extra zero modes or obstruction Euler classes are missing, the retained
zero-mode integral reduces to the incidence count \(1\), and the retained
coefficient is \(q_\Lambda\).  If the determinant-line orientation is reversed,
the coefficient changes sign; if an extra residual fermion zero mode is
present without an insertion, or if an obstruction factor is omitted, the
Berezin coefficient vanishes even though the geometric line count still
exists.

This is the path-integral bridge between the vortex calculation and the
stable-map computation.  The mirror residue computes a protected finite trace, and
the stable-map incidence chart computes a geometric count.  The A-twisted QFT
correlator is obtained only after the same regulator supplies the
vortex-normalized fugacity, the zero-mode Berezin measure, the nonzero-mode
determinant-line orientation, the compactification/contact prescription, and
the operator map to \(H\).
\end{controlledapproximation}

\begin{controlledapproximation}[Finite measure-scheme covariance]
\label{ca:cpn-degree-one-measure-scheme-covariance}
The previous bridge has a further invariance test.  The finite objects
\(c_{i,\Lambda}\), \(\dd\nu_{\Lambda,1}\),
\(\mathcal W_{\Lambda,1}^{\rm nz}\), and
\(\mathcal Z_{\Lambda,1}^{0}\) are not separately physical.  What can be
compared with the mirror residue is the transported product
\begin{equation}
  C_{\Lambda,1}^{\rm ret}
  =
  q_\Lambda
  \int_{\mathcal U_{\Lambda,1}}
  \Omega_{\Lambda,1},
  \qquad
  q_\Lambda=\exp(T)\prod_{i=1}^N c_{i,\Lambda},
  \qquad
  \Omega_{\Lambda,1}
  =
  \mathcal W_{\Lambda,1}^{\rm nz}
  \mathcal Z_{\Lambda,1}^{0}\dd\nu_{\Lambda,1}.
  \label{eq:cpn-degree-one-retained-measure-package}
\end{equation}
Consider a finite change of regulator coordinates and determinant
normalization in the degree-one sector.  Suppose
\[
  c'_{i,\Lambda}=r_i c_{i,\Lambda},\qquad
  T'=T-\sum_{i=1}^N\log r_i,
\]
with fixed logarithm branches, and suppose on a retained chart
\[
  \dd\nu'_{\Lambda,1}=J_\Lambda\,\dd\nu_{\Lambda,1},
  \qquad
  \mathcal W_{\Lambda,1}^{\prime\,{\rm nz}}
  \mathcal Z_{\Lambda,1}^{\prime\,0}
  =
  J_\Lambda^{-1}
  \mathcal W_{\Lambda,1}^{\rm nz}
  \mathcal Z_{\Lambda,1}^{0}
\]
as determinant-line densities in the same orientation convention.  Then
\begin{equation}
  q'_\Lambda=q_\Lambda,\qquad
  \Omega'_{\Lambda,1}=\Omega_{\Lambda,1},\qquad
  C_{\Lambda,1}^{\prime\,{\rm ret}}=C_{\Lambda,1}^{\rm ret}.
  \label{eq:cpn-degree-one-measure-scheme-invariance}
\end{equation}
If the determinant-line orientation is reversed, the retained integral changes
sign unless the virtual orientation and local-operator representative are
transported with the same sign.  The distinction is substantive: an
untransported sign, a stale FI coordinate, or a missing Jacobian changes the
degree-one coefficient while leaving the formal Hori--Vafa root sum looking
unchanged.  Thus the physical statement is not ``the mirror coordinate can
absorb \(c_i\)'' by itself.  The required statement is covariance of the whole
finite instanton-measure package
\eqref{eq:cpn-degree-one-retained-measure-package} through the residual
telescope below.
\end{controlledapproximation}

\begin{controlledapproximation}[Mirror residue cross-check for degree-one
instantons]
\label{ca:cpn-hv-residue-instanton-cross-check}
The residue formula
\eqref{eq:cpn-hv-residue-selection-rule} is a powerful protected calculation,
but it is not the same object as the regulated A-twisted path integral.  In a
shared finite regulator, write
\[
  q_\Lambda=\exp(T_\Lambda)\prod_i c_{i,\Lambda},
  \qquad
  S_1(q)=\mathcal R_{\mathbb P^{N-1}}(H^{2N-1})=q ,
\]
and let \(C_{\Lambda,1}^{A}\) denote the direct degree-one coefficient for
\(\langle H,H^{N-1},H^{N-1}\rangle\) in the same operator coordinate.  The
coordinate \(T_\Lambda\) carries the topological FI-theta sector weight, while
the \(c_{i,\Lambda}\) carry only the reduced determinant and operator-map
normalization.  The mirror residue may be used as a check only through the
double-entry comparison
\begin{equation}
  C_{\Lambda,1}^{A}
  -
  S_1(q_{\rm mir})
  =
  R_{\rm vort}
  +
  q_\Lambda\bigl(I_{\Lambda,1}-1\bigr)
  +
  (q_\Lambda-q_{\rm mir})
  +
  R_{\rm op}
  +
  R_{\rm cont},
  \label{eq:cpn-hv-residue-instanton-cross-check}
\end{equation}
where
\[
  I_{\Lambda,1}
  =
  \int_{\mathcal U_{\Lambda,1}}\Omega_{\Lambda,1}
\]
is the retained finite instanton-measure integral in
\eqref{eq:cpn-degree-one-retained-measure-package}.  In the degree-one
projective-space chart this symbol is not a free normalization.  Its direct
incidence part is the finite determinant
\[
  I_{\rm inc}
  =
  \det
  \frac{
    \partial(A_1,\ldots,A_{N-1},B_0,B_1-A_0,B_2,\ldots,B_{N-1})
  }{
    \partial(A_1,\ldots,A_{N-1},B_0,B_1,\ldots,B_{N-1})
  }
  =+1,
\]
together with the degree gate
\[
  1+(N-1)+(N-1)=2N-1
\]
and the compactification gate excluding the boundary strata where the two
point constraints collide or the hyperplane insertion contains one of them.
Thus \(I_{\Lambda,1}=1+R_{\rm meas}\) only after the determinant-line
orientation, zero-mode measure, compactification/contact convention, and
operator normalization have been transported to the same regulator.  The
mirror root sum has no access to these signs and gates by itself.
The term \(R_{\rm vort}\)
compares the true regulated vortex coefficient with the supplied
\(q_\Lambda I_{\Lambda,1}\) package; \(q_\Lambda(I_{\Lambda,1}-1)\) records
the zero-mode, determinant-line, orientation, compactification, and incidence
normalization that make the direct degree-one integral equal to the unit line
count; \(q_\Lambda-q_{\rm mir}\) is the finite FI-coordinate and branch
transport error between the mirror torus and the vortex measure; and
\(R_{\rm op},R_{\rm cont}\) are the local-operator and continuum comparison
remainders.

If
\[
  |R_{\rm vort}|\le B_{\rm vort},\quad
  |I_{\Lambda,1}-1|\le B_I,\quad
  |q_\Lambda-q_{\rm mir}|\le B_q,\quad
  |R_{\rm op}|\le B_{\rm op},\quad
  |R_{\rm cont}|\le B_{\rm cont},
\]
then
\begin{equation}
  \left|
    C_{\Lambda,1}^{A}
    -
    S_1(q_{\rm mir})
  \right|
  \le
  B_{\rm vort}
  +
  |q_\Lambda|B_I
  +
  B_q
  +
  B_{\rm op}
  +
  B_{\rm cont}.
  \label{eq:cpn-hv-residue-instanton-cross-check-bound}
\end{equation}
The off-pairing components must be checked at the same time:
\[
  \left|
    \eta_\Lambda(H\star_\Lambda H^{N-1},H^a)
  \right|
  \le B_{{\rm off},a},
  \qquad 0\le a<N-1,
\]
because the mirror root sum vanishes off the residue selection rule while the
direct GLSM can still receive operator-map or compactification contact
remainders.

This cross-check prevents three common substitutions.  A formal Hori--Vafa
root sum with \(q_{\rm mir}\) does not prove that the vortex determinant line
is oriented or that the residual fermion zero modes are saturated.  A
stable-map line count equal to one does not supply the vortex fugacity
\(q_\Lambda\).  A Coulomb-branch normalization check does not prove that the
same finite FI coordinate has been transported to the A-twisted observable.
Only the agreement of the residue, vortex-measure package, operator map, and
off-pairing bounds in one regulator gives a protected QFT comparison.
\end{controlledapproximation}

\begin{remark}[Degree-one stable-map incidence proof-obligation template]
\label{rem:cpn-degree-one-vortex-observable-proof-obligation-template}
The degree-one product coefficient is a physical A-twisted observable only after
the vortex normalization, determinant orientation, zero-mode integral,
compactification contribution, and local-operator map have been tied to the
same finite regulator.  The finite calculation in this paragraph deliberately
does not compute that regulated vortex fluctuation problem.  It treats the
reduced charge-one vortex coefficients \(c_{i,\Lambda}\) as supplied input from a
separate gauge-fixed construction in that regulator and checks only the
stable-map incidence coordinate that the comparison must match.  If a diagonal
spectral regulator has separately supplied the full gauge-matter deformation
complex, its
boundary conditions, zero-mode removal, and determinant-line orientation, one
may record the \(i\)th supplied coefficient in the bookkeeping form
\[
  c_{i,\Lambda}
  =
  Z_{i,\Lambda}^{\rm zm}
  \frac{
    \det F_{i,\Lambda}^{(1)}\det G_{i,\Lambda}^{(1)}
    \det B_{i,\Lambda}^{(0)}
  }{
    \det{}' B_{i,\Lambda}^{(1)}
    \det F_{i,\Lambda}^{(0)}
    \det G_{i,\Lambda}^{(0)}
  },
\]
where the prime removes the bosonic degree-one zero modes and
\(Z_{i,\Lambda}^{\rm zm}\) is the finite zero-mode metric factor in the chosen
orientation convention.  The symbols \(B,F,G\) here denote already-derived
finite spectral blocks of the full gauge, matter, and ghost complex; assigning
such blocks by hand is only determinant arithmetic and is not a derivation of
the vortex coefficient.  Then
\[
  q_\Lambda=\exp(T)\prod_{i=1}^N c_{i,\Lambda}
\]
is the supplied charge-one fugacity in that regulator.  This convention puts
the FI-theta sector weight in \(T\), not inside the \(c_{i,\Lambda}\).  What can
be checked without a vortex determinant calculation is the stable-map incidence
chart.  A degree-one map is written
\[
  \Phi_i([u:v])=A_i u+B_i v,
  \qquad i=0,\ldots,N-1,
\]
modulo the common complex scale.  Fix the three marked domain points
\([1:0]\), \([0:1]\), and \([1:1]\), fix the scale by \(A_0=1\), and impose the
incidence representatives
\[
  \Phi([1:0])=[1:0:\cdots:0],
  \qquad
  \Phi([0:1])=[0:1:0:\cdots:0],
  \qquad
  X_1-X_0=0\quad\hbox{at }[1:1].
\]
Equivalently,
\[
  A_j=0\ (j>0),\qquad
  B_0=0,\qquad
  B_1-A_0=0,\qquad
  B_j=0\ (j\ge2).
\]
With coordinates
\[
  (A_1,\ldots,A_{N-1},B_0,B_1,\ldots,B_{N-1})
\]
on this affine zero-mode chart, the incidence Jacobian is
\[
  \det
  \frac{
    \partial(A_1,\ldots,A_{N-1},B_0,B_1-A_0,B_2,\ldots,B_{N-1})
  }{
    \partial(A_1,\ldots,A_{N-1},B_0,B_1,\ldots,B_{N-1})
  }
  =+1 .
\]
Thus the incidence orientation of this affine stable-map chart is positive once
the stable-map virtual orientation has already been fixed.  The determinant-line
orientation of the regulated GLSM vortex fluctuation complex remains separate
input in the comparison.
The complex chart dimension is \(2N-1\), while the insertions
\((H,H^{N-1},H^{N-1})\) have complex degree
\[
  1+(N-1)+(N-1)=2N-1,
\]
so the stable-map selection degree is saturated.  The A-twisted fermion
Berezin measure, the insertion map that saturates it, and the comparison from
the GLSM vortex complex to this stable-map chart remain part of the residual
data below.  The two point constraints are distinct, and the hyperplane
\(X_1-X_0=0\) contains neither of them, so the contracted-boundary strata in
which a degree-zero component would carry both point insertions, or one point
insertion together with the hyperplane insertion, do not contribute in this
finite incidence chart.  With the operator \(H\) normalized by the unit
hyperplane equation, the retained incidence calculation gives
\begin{equation}
  C_{\Lambda,1}^{\rm inc}
  =
  q_\Lambda .
  \label{eq:cpn-degree-one-stable-map-incidence-retained}
\end{equation}
Changing the stable-map orientation, failing the GLSM zero-mode comparison,
allowing the two point constraints to collide, letting the hyperplane pass
through a point constraint, or rescaling the representative of \(H\) changes
this retained value.

Let
\[
  C_{\Lambda,1}
  =
  \eta_\Lambda(H\star_\Lambda H^{N-1},H^{N-1})
  =
  \big\langle H,H^{N-1},H^{N-1}\big\rangle_{\Lambda,d=1}^{A}
\]
be the full degree-one three-point coefficient in the same operator coordinate.
Equation \eqref{eq:cpn-degree-one-stable-map-incidence-retained} is only the
retained finite stable-map incidence chart with supplied vortex coefficients.
A usable QFT comparison therefore has the conditional
form
\begin{equation}
  C_{\Lambda,1}-q_\Lambda
  =
  R_{\rm coeff}
  +R_{\rm det}
  +R_{\rm zm}
  +R_{\rm line}
  +R_{\rm op}
  +R_{\rm cont},
  \label{eq:cpn-degree-one-vortex-observable-telescope}
\end{equation}
where \(R_{\rm coeff}\) compares the retained charge-one vortex integral with
the coefficients \(c_{i,\Lambda}\), \(R_{\rm det}\) tracks determinant-line and
orientation choices, \(R_{\rm zm}\) tracks the universal and residual fermion
zero modes, \(R_{\rm line}\) tracks compactification, obstruction, and boundary
corrections to the degree-one line count, \(R_{\rm op}\) tracks the local
operator map to the hyperplane class, and \(R_{\rm cont}\) is the continuum
comparison remainder.  If \(|R_\alpha|\le\epsilon_\alpha\), then
\begin{equation}
  |C_{\Lambda,1}-q_\Lambda|
  \le
  \epsilon_{\rm coeff}
  +\epsilon_{\rm det}
  +\epsilon_{\rm zm}
  +\epsilon_{\rm line}
  +\epsilon_{\rm op}
  +\epsilon_{\rm cont}.
  \label{eq:cpn-degree-one-vortex-observable-bound}
\end{equation}
The same stable-map degree count gives no retained off-pairing for
\(0\le a<N-1\), but the full shared-regulator statement still requires the
separate bounds
\begin{equation}
  \left|
    \eta_\Lambda(H\star_\Lambda H^{N-1},H^a)
  \right|
  \le \epsilon_{{\rm off},a}.
  \label{eq:cpn-degree-one-off-pairing-bound}
\end{equation}
Equations
\eqref{eq:cpn-degree-one-vortex-observable-bound}--\eqref{eq:cpn-degree-one-off-pairing-bound}
are a proof-obligation template for the product
\[
  H\star H^{N-1}=q_{\rm phys}\,1.
\]
They become evidence for the QFT correlator only when the coefficient,
determinant, zero-mode, compactification, operator-map, off-pairing, and
continuum residuals are supplied by independent estimates in the shared
regulator.  The mirror residue gives the same algebraic root sum, and the
stable-map dimension count gives the geometric line count; neither replaces the
determinant and fluctuation comparison in the A-twisted path integral.
\end{remark}

\begin{controlledapproximation}[Degree-\(d\) projective instanton iteration]
\label{ca:cpn-degree-d-vortex-observable-iteration}
For \(X=\mathbb P^{N-1}\), the mirror residue calculation predicts
\[
  \mathcal R_\Lambda(H^{N-1+dN})=q_\Lambda^d,\qquad d\ge0,
\]
in the same coordinate in which the degree-one product reads
\(H^N=q_\Lambda\).  The exponent \(d\) carries the topological sector
weight: the K\"ahler area and flat \(B\)-field phase enter through
\(q_\Lambda^d\).  This compact formula should be read as a sharp protected
prediction, not as a substitute for the degree-\(d\) instanton computation.
A direct A-twisted path integral must still isolate the degree-\(d\) sector,
construct its collective-coordinate measure, remove the universal fermion
zero modes, saturate the residual Berezin measure by the chosen observable,
transport the nonzero-mode determinant line and orientation, control boundary
strata in the stable-map compactification, and compare the operator
representative with the GLSM observable.

Let
\[
  T_{\Lambda,d}
  =
  \eta_\Lambda\!\left(
    H^{N-1}\star_\Lambda (H^N)^{\star d},1
  \right),
  \qquad T_{\Lambda,0}=1,
\]
where \(\eta_\Lambda\) is the A-twisted topological metric and the
normalization of \(H\) is the one used in
Remark~\ref{rem:cpn-degree-one-vortex-observable-proof-obligation-template}.
Equivalently, one may distribute the same total hyperplane power among marked
local operators before using the chosen contact-term convention for the
quantum product.  If the degree-one product is used as the elementary gluing
step, the degree-\(d\) comparison has the conditional telescope
\begin{equation}
  T_{\Lambda,d}-q_\Lambda^d
  =
  \sum_{j=1}^{d} q_\Lambda^{d-j}
    \big(R_{{\rm gl},j}+R_{{\rm off},j}\big)
  +R_{{\rm det},d}
  +R_{{\rm zm},d}
  +R_{{\rm cpt},d}
  +R_{{\rm op},d}
  +R_{{\rm cont},d}.
  \label{eq:cpn-degree-d-vortex-observable-telescope}
\end{equation}
Here \(R_{{\rm gl},j}\) is the error in replacing the \(j\)-th sector
assembly by a glued degree-one block, \(R_{{\rm off},j}\) measures leakage
into lower pairings during the iteration, \(R_{{\rm det},d}\) and
\(R_{{\rm zm},d}\) are the determinant-line and fermion-measure comparisons
in the degree-\(d\) sector, \(R_{{\rm cpt},d}\) records boundary and
obstruction corrections, \(R_{{\rm op},d}\) records the local representative
of \(H^{N-1+dN}\), and \(R_{{\rm cont},d}\) is the continuum-regulator
remainder.  Thus the usable bound is
\begin{equation}
  |T_{\Lambda,d}-q_\Lambda^d|
  \le
  \sum_{j=1}^{d}|q_\Lambda|^{d-j}
    \big(\epsilon_{{\rm gl},j}+\epsilon_{{\rm off},j}\big)
  +\epsilon_{{\rm det},d}
  +\epsilon_{{\rm zm},d}
  +\epsilon_{{\rm cpt},d}
  +\epsilon_{{\rm op},d}
  +\epsilon_{{\rm cont},d}.
  \label{eq:cpn-degree-d-vortex-observable-bound}
\end{equation}
For \(d=1\) this specializes to the proof-obligation map above.  For
\(d>1\), a product of degree-one incidence counts supplies only the expected
enumerative skeleton.  It does not fix the determinant line on the
multi-instanton boundary, the contact terms when marked points collide, the
orientation of glued obstruction factors, or the zero-mode pairing that kills
off-channel contributions.  Conversely, the mirror root sum is a stringent
test of the direct construction: once the sector measure, gluing, operator
normalization, and contact terms are transported to the same coordinate, every
term on the right side of
\eqref{eq:cpn-degree-d-vortex-observable-telescope} must vanish or be bounded
well enough to recover \(q_\Lambda^d\).
\end{controlledapproximation}

\section{The Cigar--Liouville Mirror Chain as a QFT Duality Problem}
\label{sec:cigar-liouville-mirror-qft-problem}

The same dualization mechanism gives a useful intrinsic two-dimensional
description of the relation between the \(\mathcal N=2\) cigar coset and
\(\mathcal N=2\) Liouville theory.  This subsection records the construction
as a QFT comparison problem.  The names ``cigar'' and ``Liouville'' refer to
two-dimensional field theories and their semiclassical target descriptions.

\begin{conjecture}[Cigar--Liouville mirror QFT equivalence]
\label{conj:cigar-liouville-mirror-qft}
Fix a level \(k>0\) in the conventions used below, a spin structure, \(R\)-
and flavor-background fields, contact-term conventions, and a continuum
definition of the axial \(\mathcal N=2\) cigar coset
\(SL(2,\mathbb R)_k/U(1)\) or of the logarithmic-chiral GLSM flowing to it.
Fix also a regulated \(\mathcal N=2\) Liouville theory with the corresponding
linear-dilaton/background-charge convention, periodic chiral or twisted-chiral
coordinate, exponential interaction, \(D\)-term/Kahler normalization,
functional measure, noncompact boundary conditions at the Liouville wall and
weak-coupling end, and the same background data.

The conjecture asserts an equivalence of the infrared endpoints of these
regulated theories.  At the endpoint the Hamiltonians and \(\mathcal N=2\)
superconformal generators agree under a unitary map of Hilbert spaces;
spectral-flow sectors, spin structures, and discrete and continuous
representations are matched; normalizable states, nonnormalizable source
operators, reflection amplitudes, local OPEs, partition functions, and
correlation functions are identified with their contact terms; and the
equivalence respects topological twists, defects, branes, and boundary states.
The semiclassical quotient metric, the asymptotic dualization, the single
exponential from the ordinary charged chiral multiplet, and the absence of a
smooth logarithmic-chiral vortex exponential are evidence for this statement.
They leave separate continuum obligations: the continuous-spectrum measure,
operator completeness, strong-coupling control at the cigar tip or Liouville
wall, finite-field \(D\)-term/Kahler control, classification of allowed
symmetry-preserving deformations, and the defect/boundary matching.
\end{conjecture}

For Conjecture~\ref{conj:cigar-liouville-mirror-qft}, the checks in this
section form the following protected-data and proof-debt ledger:
\[
  \begin{array}{rcl}
    C_{\rm cl} &:& \hbox{classical quotient metric and asymptotic cylinder},\\
    C_{\rm dual} &:& \hbox{periodic dual variables and vector constraint},\\
    C_{\rm vort} &:& \hbox{ordinary vortex exponential and logarithmic-vortex obstruction},\\
    C_{\rm spec} &:& \hbox{spectral flow, continuous measure, and reflection data},\\
    C_{\rm bdry} &:& \hbox{defect, brane, boundary-state, and contact-term matching}.
  \end{array}
\]
The first three entries are protected construction data.  The spectral entries
below record the exact normalization targets that a full proof must hit; they
are not consequences of the superpotential alone.
For the boundary entry, the relevant targets are the annulus amplitudes,
boundary-state one-point functions, defect-twined traces, and open-channel
multiplicities of
Controlled~Approximation~\ref{ca:glsm-mirror-boundary-defect-gate}, specialized
to the cigar/Liouville Hilbert space and spin structures.

Start with a \(U(1)\) vector multiplet \(V\), an ordinary chiral multiplet
\(\Phi\) of charge \(1\), and a logarithmic chiral multiplet
\[
  P\sim P+2\pi\ii
\]
whose imaginary part shifts under the gauge symmetry so that
\[
  P+\overline P+V
\]
is gauge invariant.  With \(k>0\), the bosonic low-derivative part of the
GLSM contains
\begin{equation}
  |(\partial_\mu-\ii A_\mu)\phi|^2
  +
  \frac{k}{2}(\partial_\mu\operatorname{Re}p)^2
  +
  \frac{k}{2}
  (\partial_\mu\operatorname{Im}p-A_\mu)^2
  \label{eq:cigar-glsm-bosonic-kinetic}
\end{equation}
together with the auxiliary constraint
\[
  |\phi|^2+\frac{k}{2}(p+\overline p)=0
\]
in the low-energy quotient.  Write
\[
  \rho=\sqrt2\,|\phi|,\qquad \vartheta=\arg\phi,\qquad
  p=u+\ii\chi .
\]
Then
\[
  |(\partial_\mu-\ii A_\mu)\phi|^2
  =
  \frac12(\partial_\mu\rho)^2
  +
  \frac{\rho^2}{2}(\partial_\mu\vartheta-A_\mu)^2 .
\]
The auxiliary constraint gives \(u=-\rho^2/(2k)\), hence
\[
  \frac{k}{2}(\partial_\mu u)^2
  =
  \frac{\rho^2}{2k}(\partial_\mu\rho)^2 .
\]
Fix the gauge \(\chi=0\).  The \(A_\mu\)-dependent angular part is
\[
  \frac12\rho^2(\partial_\mu\vartheta-A_\mu)^2
  +
  \frac{k}{2}A_\mu A^\mu .
\]
Its algebraic equation is
\begin{equation}
  (\rho^2+k)A_\mu=\rho^2\partial_\mu\vartheta .
  \label{eq:cigar-glsm-gauge-field-saddle}
\end{equation}
Substituting
\[
  A_\mu=\frac{\rho^2}{\rho^2+k}\partial_\mu\vartheta
\]
gives the angular coefficient
\[
  \frac12
  \frac{k\rho^2}{\rho^2+k}
  (\partial_\mu\vartheta)^2
  =
  \frac12
  \frac{\rho^2}{1+\rho^2/k}
  (\partial_\mu\vartheta)^2 .
\]
Combining the radial and angular terms, the classical two-real-dimensional
quotient metric is
\begin{equation}
  ds^2
  =
  \left(1+\frac{\rho^2}{k}\right)d\rho^2
  +
  \frac{\rho^2}{1+\rho^2/k}\,d\vartheta^2 .
  \label{eq:cigar-glsm-classical-quotient-metric}
\end{equation}
This metric is the classical quotient metric.  The conformal cigar coset is
an infrared fixed-point statement whose definition requires the corresponding
quantum current algebra or continuum fixed-point construction, rather than
the low-derivative quotient metric alone.

Dualize both logarithmic directions.  Introduce twisted chiral variables
\[
  Y,\qquad Y_P,
  \qquad
  Y\sim Y+2\pi\ii,\quad Y_P\sim Y_P+2\pi\ii .
\]
The ordinary charged chiral multiplet supplies the exponential term
\(\mu e^{-Y}\).  The logarithmic chiral multiplet contributes a quadratic
dual Kahler term rather than a second exponential term.  The absence of a
\(P\)-vortex exponential has a finite topological reason in the smooth
finite-action sector.  The periodic field \(e^{\ii\chi}\) is a unit-norm
charged field for the same \(U(1)\) connection \(A\):
\[
  D(e^{\ii\chi})
  =
  \ii e^{\ii\chi}(d\chi-A).
\]
Thus a smooth \(P\)-configuration on a Euclidean surface \(\Sigma_E\)
supplies a nowhere-vanishing section of the associated unit circle bundle.
If the gauge bundle has vortex number
\[
  n=\frac{1}{2\pi}\int_{\Sigma_E}F\in\mathbb Z ,
\]
such a nowhere-vanishing section trivializes the line bundle and forces
\(n=0\).  Equivalently, on a disk, a boundary winding of \(\chi\) extends
smoothly through the center only when the winding is zero; unlike the
ordinary charged scalar \(\phi=\rho e^{\ii\vartheta}/\sqrt2\), the
logarithmic multiplet has no radial modulus whose zero can absorb the
winding.  Allowing singular \(P\)-fields would be a different defect
insertion, not the smooth finite-action vortex sector used to generate
twisted \(F\)-terms.

The dual \(D\)-term before the vector constraint is just as important as the
twisted superpotential.  Eliminating the real first-order fields gives, up to
affine \(D\)-terms and the common normalization convention of the previous
section,
\begin{equation}
  S_{\rm dual}^{D}
  =
  -\frac1{4\pi}
  \int d^2z\,d^4\theta
  \left[
  -\frac1{g^2}\overline\Sigma\Sigma
  -\frac12(Y+\overline Y)\log(Y+\overline Y)
  -\frac1{2k}\overline Y_PY_P
  \right]
  +
  S_{\Sigma(Y+Y_P)} .
  \label{eq:cigar-liouville-intermediate-dual-dterm}
\end{equation}
The term \(S_{\Sigma(Y+Y_P)}\) is the twisted \(F\)-term linear in
\(\Sigma(Y+Y_P)\) displayed below.  Equation
\eqref{eq:cigar-liouville-intermediate-dual-dterm} records the full-action
data that the exponential alone cannot provide: the ordinary charged chiral
again produces the singular half-cylinder metric in \(Y+\overline Y>0\), while
the logarithmic chiral produces a quadratic dual \(D\)-term for \(Y_P\).  After
the vector constraint \(Y_P=-Y\), quantum corrections replace the displayed
classical expression by an effective \(K(Y,\overline Y)\).  The finite-field
part of this \(K\) is not determined by the protected twisted \(F\)-term.

In the convention compatible with
\eqref{eq:glsm-hv-dual-twisted-superpotential}, the
low-energy twisted superpotential has the form
\begin{equation}
  \widetilde W_{\rm cig}(\Sigma,Y,Y_P)
  =
  -\Sigma(Y+Y_P)
  -
  \mu e^{-Y},
  \label{eq:cigar-liouville-dual-superpotential-before-constraint}
\end{equation}
up to affine redefinitions of \(Y_P\) and \(\Sigma\).  The vector multiplet
imposes
\begin{equation}
  Y+Y_P=0 .
  \label{eq:cigar-liouville-linear-constraint}
\end{equation}
Substituting the constraint leaves one twisted chiral variable \(Y\) with
twisted superpotential
\begin{equation}
  \widetilde W_{\rm Liouville}(Y)
  =
  -\mu e^{-Y}.
  \label{eq:n2-liouville-twisted-superpotential}
\end{equation}
The dual Kahler potential has the asymptotic large-\(\operatorname{Re}Y\)
behavior
\begin{equation}
  K(Y,\overline Y)
  =
  \frac{1}{8\pi k}\,\overline Y Y
  +
  O(e^{-\operatorname{Re}Y}),
  \qquad
  \operatorname{Re}Y\to+\infty,
  \label{eq:n2-liouville-asymptotic-kahler}
\end{equation}
with the sign convention appropriate to twisted chiral superspace.  At this
stage \(Y\) is still a twisted chiral field.  Thus
\eqref{eq:n2-liouville-twisted-superpotential} is the twisted-chiral
Liouville presentation produced by the GLSM dualization, not yet the ordinary
chiral Liouville action often used to describe the endpoint SCFT.  Passing to
that endpoint presentation requires a mirror/T-duality automorphism that maps
the twisted \(F\)-term measure to the ordinary chiral \(F\)-term measure, sends
the periodic scalar and fermions to their Liouville variables, and transports
the \(R\)-current, supercharges, and boundary conditions at the noncompact end.
The rescaling \(Y=\sqrt{k}\mathcal Y\) below fixes normalization after this map;
it is not by itself the chirality-changing step.

The full cigar/Liouville comparison must also hit the exact spectral
normalization.  In the coset convention compatible with the asymptotic
cylinder, a standard large-level sigma-model representative is
\begin{equation}
  ds^2_{\rm cig}
  =
  k\left(dr^2+\tanh^2 r\,d\chi^2\right),
  \qquad
  \Phi_{\rm cig}=-\log\cosh r,
  \qquad
  \chi\sim\chi+2\pi ,
  \label{eq:n2-cigar-metric-dilaton-representative}
\end{equation}
so \(r\to\infty\) is a circle of radius \(\sqrt{k}\) together with a linear
dilaton of background charge \(Q=1/\sqrt{k}\).  This metric/dilaton pair is
not the finite-level QFT datum by itself: away from the asymptotic region it is
scheme- and field-redefinition-dependent, and the additive dilaton constant
fixes the normalization of noncompact amplitudes.  The exact finite-level
object is the noncompact supercoset with its Hilbert space, branching data,
reflection relation, normalizability prescription, spin structures, and sewing
constraints.  Its central charge is
\begin{equation}
  c_{\rm cig}=c_{\rm Liouville}=3+\frac6k .
  \label{eq:n2-cigar-liouville-central-charge}
\end{equation}
The large value of \(c_{\rm cig}\) as \(k\to0\) is not a compact
\(c\)-theorem paradox.  The noncompact theory has no normalizable conformally
invariant vacuum; the bottom of the continuous spectrum in the zero
momentum/winding sector has
\[
  h_{\min}=\frac1{4k},
  \qquad
  c_{\rm eff}=c_{\rm cig}-24h_{\min}=3,
\]
so finite-size state counting and vacuum normalizability are part of the
comparison data.

The continuous and discrete sectors are not visible in
\(\widetilde W_{\rm Liouville}\) alone.  In the Volume~V convention the
supersymmetric \(SL(2,\mathbb R)_k/U(1)\) coset is built from bosonic
\(\widehat{\mathfrak{sl}}(2,\mathbb R)_{k+2}\) primaries and two charged
fermions.  A coset primary is labelled by \((j,m,\overline m)\), with
\[
  h=\frac{-j(j-1)+m^2}{k},
  \qquad
  q=\frac{2m}{k},
  \qquad
  \overline h=\frac{-j(j-1)+\overline m^2}{k},
  \qquad
  \overline q=\frac{2\overline m}{k}.
\]
The continuous series has \(j=\frac12+\ii s\), \(s\in\mathbb R_{\ge0}\),
with \(m-\overline m\in\mathbb Z\) and \(m+\overline m\in k\mathbb Z\);
the discrete series lies at
\[
  \frac12<j<\frac{k+1}{2}
\]
with \(m,\overline m\) in the corresponding highest- or lowest-weight towers.
Spectral flow by \((\eta,\overline\eta)\in\frac12\mathbb Z^2\) sends
\[
  h\mapsto
  \frac{-j(j-1)+(m+\eta)^2}{k}+\frac{\eta^2}{2},
  \qquad
  q\mapsto\frac{2(m+\eta)}{k}+\eta,
\]
and similarly on the right.  In the asymptotic circle variables the momentum
and winding are
\begin{equation}
  n=m+\eta-\overline m-\overline\eta,
  \qquad
  w=\frac{m+\eta+\overline m+\overline\eta}{k},
  \label{eq:n2-cigar-spectral-flow-momentum-winding}
\end{equation}
so the flowed sector condition is \(w\in\mathbb Z\).
The noncompact field identification extending the discrete range is
\begin{equation}
  \Psi^{sl,(\eta,\overline\eta)}_{j,m,\overline m}
  \simeq
  \Psi^{sl,(\eta+1,\overline\eta+1)}
       _{\frac{k+2}{2}-j,\,
         m-\frac{k+2}{2},\,\overline m-\frac{k+2}{2}} .
  \label{eq:n2-cigar-noncompact-field-identification}
\end{equation}
It preserves the exact circle momentum \(n\) and changes the winding
representative by an integral unit.  There is no independent identification
with opposite left/right spectral-flow shifts: that shortcut would change the
exact momentum sector.  The spin-structure choice and modular left-right
pairing determine which flowed sectors occur in the Hilbert space and must be
matched by the Liouville presentation.

The reflection amplitude is part of the operator normalization.  With
\(\delta=2j-1\),
\begin{equation}
  R(j,m,\overline m;k)
  =
  \nu(k)^{\delta}
  \frac{
  \Gamma(1-\delta/k)\Gamma(j+m)\Gamma(j-\overline m)\Gamma(-\delta)}
  {\Gamma(1+\delta/k)\Gamma(m-j+1)
  \Gamma(-\overline m-j+1)\Gamma(\delta)},
  \qquad
  \nu(k)=\frac1\pi
  \frac{\Gamma(1+1/k)}{\Gamma(1-1/k)} .
  \label{eq:n2-cigar-reflection-amplitude}
\end{equation}
For \(j=\frac12+\ii s\), \(s\in\mathbb R\), and generic positive \(k\) with the
positive real branch of \(\nu(k)\), the gamma-function expression gives
\[
  R(1-j,m,\overline m;k)=R(j,m,\overline m;k)^{-1},
  \qquad |R(j,m,\overline m;k)|=1
\]
in the continuous series.  The reflection target therefore carries dynamical
spectral data: its phase contributes to the radial density of states by
\[
  \rho_{\rm refl}(s;m,\overline m)
  =
  \frac1{2\pi}\,\partial_s
  \arg R\!\left(\frac12+\ii s,m,\overline m;k\right).
\]
For example, the factor \(\nu(k)^{2\ii s}\) changes this density by
\(\pi^{-1}\log\nu(k)\).  Omitting it can preserve \(|R|=1\) while changing the
continuum normalization.  The poles similarly carry operator data.  If
\(j_0+m=-n\), \(n\in\mathbb Z_{\ge0}\), is a simple pole not colliding with
another gamma factor, then with \(\delta_0=2j_0-1\)
\[
  \operatorname*{Res}_{j=j_0} R(j,m,\overline m;k)
  =
  \nu(k)^{\delta_0}
  \frac{
  \Gamma(1-\delta_0/k)\Gamma(j_0-\overline m)\Gamma(-\delta_0)}
  {\Gamma(1+\delta_0/k)\Gamma(m-j_0+1)
  \Gamma(-\overline m-j_0+1)\Gamma(\delta_0)}
  \frac{(-1)^n}{n!}.
\]
Analogous residues come from the other gamma factors and must match the
discrete-state normalization on the Liouville side.

In this chapter the displayed formula is still an imported normalization
target, not yet a derivation from the Liouville path integral.  The constant
\(\nu(k)\), the Liouville coupling \(\mu\), and the additive dilaton constant
must be fixed in the same scheme; special levels where the gamma factors have
zeros or poles, for instance \(k=1\), require a limiting or renormalized
operator prescription before the preceding density and residue formulae are
used.  The full continuous spectral measure, all pole residues,
spin-structure pairing, and the reflection relation
\[
  \Psi_{j,m,\overline m}
  =
  R(j,m,\overline m;k)\Psi_{1-j,m,\overline m}
\]
are therefore \(C_{\rm spec}\) data.  A candidate Liouville action with the
same exponential but a different finite-field \(K(Y,\overline Y)\) is a
different theory unless this spectral package and the allowed deformation
classification agree.

There is a concrete spectral diagnostic behind this warning.  After fixing the
asymptotic circle quantum numbers, a continuous radial channel is a
self-adjoint one-dimensional problem on a noncompact end, in the form
\[
  \left[
    -\partial_\varrho^2
    +\frac{Q^2}{4}
    +V_{\rm wall}(\varrho)
    +V_{K}(\varrho)
  \right]\psi_p
  =
  \left(p^2+\frac{Q^2}{4}\right)\psi_p .
\]
The exponential wall supplies one part of the potential, but the finite-field
Kahler/measure data determine both \(V_K\) and the self-adjoint boundary
condition at the strong-coupling end.  Even the toy channel with
\(V_{\rm wall}=V_K=0\) on \(\varrho\ge0\) and Robin condition
\(\psi'(0)=\lambda\psi(0)\) has reflection amplitude
\[
  R_\lambda(p)
  =
  \frac{\lambda+\ii p}{\ii p-\lambda},
  \qquad
  |R_\lambda(p)|=1,
\]
and changes the continuum density by
\[
  \rho_\lambda(p)-\rho_0(p)
  =
  \frac1{2\pi}\partial_p\arg R_\lambda(p)
  =
  \frac{\lambda}{\pi(\lambda^2+p^2)} .
\]
Thus the asymptotic central charge, the exponential \(F\)-term, and unitarity
of the reflection coefficient do not fix the spectral measure.  A total
radial derivative in a \(D\)-term has the same obstruction:
\(\int_0^\infty \partial_\varrho J\,d\varrho=J(\infty)-J(0)\), so it vanishes
only after the boundary current at the Liouville wall is controlled or included
as a contact/source term.  A full cigar/Liouville proof must therefore show
that the actual finite-field \(K\), measure, and wall boundary condition flow
to the same \(R(p)\) and pole residues, with the protected superpotential only
one entry in the datum.

The same point appears in a finite-volume Hamiltonian regulator.  Put the
radial channel on \(0\le\varrho\le L\), impose
\(\psi(L)=0\) at the far wall, and keep the same Robin condition at the
Liouville wall.  Then the momenta solve
\begin{equation}
  -p\cos(pL)=\lambda\sin(pL),
  \qquad
  E_n(L)=p_n(L)^2+\frac{Q^2}{4}.
  \label{eq:liouville-robin-finite-volume-quantization}
\end{equation}
For \(\lambda=0\) this gives \(p_nL=(n+\tfrac12)\pi\), while a nonzero
\(\lambda\) shifts every root before the limit \(L\to\infty\).  Equivalently,
the large-\(L\) root-counting function contains
\[
  N_\lambda(p;L)
  =
  \frac{Lp}{\pi}
  +\frac1{2\pi}\arg R_\lambda(p)+O(L^{-1}).
\]
Thus the boundary functional changes the regulated Hamiltonian spectrum and
then survives as the continuum phase density.  A mirror proposal that does not
specify this self-adjoint domain has not specified the Hilbert space whose
equality is being conjectured.

\begin{construction}[Source spectral resolution]
\label{constr:cigar-liouville-source-spectral-resolution}
Fix one cigar/Liouville continuous-series channel, including the circle
momentum, winding representative, spin structure, and left-right pairing.
Let \(I\subset\mathbb R_{\ge0}\) be a compact momentum window with no pole
crossing in the reflection normalization.  The finite-box roots are tested
against the phase condition
\[
  2sL+\arg R\!\left(\frac12+\ii s,m,\overline m;k\right)
  =(2n+1)\pi ,
\]
up to the far-wall convention.  The continuum object compared by a local
source \(O\) is not the phase alone, but the source-normalized spectral
measure
\begin{equation}
  d\mu_{O,\alpha}(s)
  =
  |A_{O,\alpha}(s)|^2
  \left[
    d\mu^{(0)}_{\alpha}(s)
    +\frac1{2\pi}\partial_s
      \arg R\!\left(\frac12+\ii s,m,\overline m;k\right)ds
  \right]
  +\sum_{b\in{\mathcal P}_\alpha}
    |A_{O,\alpha,b}|^2\,{\mathcal R}_{\alpha,b}\,\delta(s-s_b).
  \label{eq:cigar-liouville-source-spectral-measure}
\end{equation}
Here \(\alpha\) abbreviates the channel labels,
\(d\mu^{(0)}_{\alpha}\) is the declared reference Plancherel measure in the
same operator normalization, \(A_{O,\alpha}(s)\) is the continuum source row,
and \({\mathcal R}_{\alpha,b}\) is the residue normalization of a discrete or
resonant pole included in the chosen contour prescription.  In a finite-box
regulator the matching condition is that, for every smooth test \(w\) supported
in \(I\),
\[
  \sum_{s_{n,L}\in I}
    w(s_{n,L})\,|A^{(L)}_{O,\alpha,n}|^2
  \longrightarrow
  \int_I w(s)\,d\mu_{O,\alpha}(s)
\]
after the reference volume normalization and contact counterterms have been
fixed.  The Euclidean source two-point function in the same channel is then
\begin{equation}
  G_{O,\alpha}^{I}(\beta)
  =
  \int_I
    \exp\!\left[-\beta\left(s^2+\frac{Q^2}{4}+E_\alpha^\perp\right)\right]
    d\mu_{O,\alpha}(s)
  +C_{O,\alpha}^{\rm ct}(\beta;I),
  \label{eq:cigar-liouville-source-spectral-two-point}
\end{equation}
with \(E_\alpha^\perp\) the fixed angular/fermionic contribution and
\(C^{\rm ct}\) the local contact piece.  Therefore a mirror endpoint with the
same central charge, asymptotic Liouville exponential, and unitary reflection
coefficient can still fail the QFT comparison if it changes
\(\partial_s\arg R\), the source row \(A_{O,\alpha}\), a pole residue, or the
contact term.  This is the first observable-level use of \(C_{\rm spec}\):
reflection data, Plancherel normalization, residues, and operator sources must
assemble into the same \(G_O(\beta)\), with the protected superpotential only
one member of the comparison datum.
\end{construction}

In the asymptotic Liouville target variables write
\[
  Y=\sqrt{k}\,{\mathcal Y},
  \qquad
  {\mathcal Y}=\varrho+\ii\widetilde y+\hbox{fermions},
  \qquad
  \widetilde y\sim\widetilde y+\frac{2\pi}{\sqrt{k}} .
\]
After the mirror/T-duality automorphism described above has been supplied, the
ordinary chiral endpoint presentation has the flat-worldsheet asymptotic
superspace form
\begin{align}
  S_L^{\rm flat}
  &=
  \int d^2z
  \left[
  -\frac1{8\pi}\int d^4\theta\,{\mathcal Y}\overline{\mathcal Y}
  +
  \mu\int d^2\theta\,e^{-\sqrt{k}{\mathcal Y}}
  +
  \overline\mu\int d^2\overline\theta\,
  e^{-\sqrt{k}\overline{\mathcal Y}}
  \right],
  \label{eq:n2-liouville-normalized-action}\\
  S_L^{\rm bg}
  &=
  \frac{Q}{8\pi}\int d^2z\,\sqrt h\,\mathcal R^{(2)}\varrho,
  \qquad
  Q=\frac1{\sqrt{k}} .
  \label{eq:n2-liouville-background-charge}
\end{align}
The flat superspace action alone hides the background charge.  In the
asymptotic convention used here, the curvature coupling fixes the improvement
terms in the generators:
\[
  T_{\rm asymp}=T_{\rm free}-Q\,\partial^2\varrho,
  \qquad
  J_R=J_R^{\rm ferm}+2\ii Q\,\partial\widetilde y,
\]
with the corresponding derivative terms in \(G^\pm\).  The exponential has
left and right weight \(1/2\) before the two supercharges act: the radial
factor contributes \(1/2-k/4\) and the dual-circle factor contributes
\(k/4\).  Equivalently, its component wall contains the asymptotic interaction
\[
  \Delta S_{\rm wall}
  =
  \int d^2z\,
  \mu\bigl(k\lambda^+\widetilde\lambda^+
  +\sqrt{k}F\bigr)e^{-\sqrt{k}(\varrho+\ii\widetilde y)}
  +\text{c.c.}
\]
This is why the integrated \(G_{-1/2}^+\overline G_{-1/2}^+\) deformation is
marginal in the linear-dilaton normalization.

The Hori--Kapustin \(\kappa\)-deformation route is therefore best read here as
a physical-continuity argument unless each global hypothesis is supplied.  One
needs existence of the fixed point along the \(\kappa\)-path, continuity of the
noncompact spectrum and boundary conditions, absence of bifurcation or phase
transition, local rigidity against normalizable marginal deformations, and
control of nonnormalizable boundary deformations.  Local rigidity alone does
not exclude a disconnected fake-Liouville fixed point with the same asymptotic
superpotential.

\begin{controlledapproximation}[Pathwise fake-Liouville exclusion gate]
\label{ca:cigar-liouville-pathwise-fake-fixed-point-gate}
To turn the preceding continuity argument into a QFT statement, one must
transport more than protected endpoint labels.  Put the radial direction in a
box of length \(L\), keep a UV cutoff \(\Lambda\), and let
\(\kappa\in[0,1]\) parametrize a proposed deformation from the weakly
controlled presentation to the desired Liouville endpoint.  In a fixed
energy/source window the transported datum has the form
\[
  \begin{aligned}
  {\mathcal S}_{\Lambda,L}(\kappa)
  =
  \bigl(
  &\{E_{n,\Lambda,L},P_{n,\Lambda,L}\}_{n\le N},
    \{\theta_{\Lambda,L}(s_\ell)\}_{\ell\le M},
    \{\operatorname*{Res}_{j_a}R_{\kappa}\}_{a\le A},
  \\
  &\{Z_{\alpha\beta,\Lambda,L}\}_{\alpha,\beta\le B},
    \{{\mathcal O}_{A,\Lambda,L}e_{0,\Lambda,L}\}_{A\le C},
    B_{\infty,\Lambda,L}
  \bigr)(\kappa).
  \end{aligned}
\]
where \(P_n\) are spectral projectors, \(\theta=\arg R\) is the reflection
phase in the same normalization as \eqref{eq:n2-cigar-reflection-amplitude},
\(Z_{\alpha\beta}\) are boundary cylinder amplitudes, and the last two entries
record source vectors and the noncompact wall domain.  A pathwise proof must
provide a matching of these entries and a residual estimate such as
\begin{equation}
  \begin{aligned}
  &\|P_n(\kappa')-P_n(\kappa)\|
  +|E_n(\kappa')-E_n(\kappa)|
  +|\theta(s_\ell;\kappa')-\theta(s_\ell;\kappa)|\\
  &\quad
  +|\operatorname*{Res}_{j_a}R_{\kappa'}
       -\operatorname*{Res}_{j_a}R_{\kappa}|
  +|Z_{\alpha\beta}(\kappa')-Z_{\alpha\beta}(\kappa)|
  \le C|\kappa'-\kappa|+\varepsilon_{\Lambda,L}
  \end{aligned}
  \label{eq:cigar-liouville-pathwise-spectral-transport}
\end{equation}
on each chart, with \(\varepsilon_{\Lambda,L}\to0\) in the continuum limit,
and with no resonance pole crossing the chosen contours, no boundary condition
leaving the self-adjoint domain, and no source vector leaving the operator
topology being compared.

This gate is stronger than local deformation rigidity.  The protected package
\[
  (c,\ Q,\ \widetilde W_{\rm Liouville},\
    K_{\rm asymp},\ H^1_{\rm def}=0)
\]
can agree at two isolated endpoints while
\({\mathcal S}_{\Lambda,L}\) differs in a reflection phase, a pole residue, a
boundary annulus amplitude, or a source matrix element.  Such a theory is the
``fake Liouville'' possibility.  It is excluded by a pathwise spectral,
boundary, and source transport theorem, not by the absence of normalizable
marginal tangent vectors at one endpoint.
\end{controlledapproximation}

The construction supplies a precise protected-data bridge:
\[
  \begin{aligned}
  &\text{charged chiral dualization}
  +\text{nonzero vortex exponential}
  +\text{IR vector constraint}
  \\
  &\hspace{3em}\Longrightarrow
  \mathcal N=2\ \text{protected Liouville \(F\)-term/asymptotics}.
  \end{aligned}
\]
This is the \(C_{\rm cl}\), \(C_{\rm dual}\), and \(C_{\rm vort}\) evidence
lane for Conjecture~\ref{conj:cigar-liouville-mirror-qft}, supplemented by a
\(C_{\rm spec}\) normalization target.  It identifies the asymptotic variables,
the protected Liouville superpotential, the exact coset spectral labels, and
the reflection-amplitude data that the infrared equivalence must reproduce.
It still does not prove equality of continuum Hilbert spaces, local algebras,
\(\mathcal N=2\) generators, defect categories, full continuous-spectrum
measure, pole-residue normalization, or boundary states.  Those remain
\(C_{\rm spec}\) and \(C_{\rm bdry}\) obligations, together with the rigidity
statement excluding a different finite-field \(K(Y,\overline Y)\) with the
same asymptotics and superpotential.  The semiclassical quotient metric
\eqref{eq:cigar-glsm-classical-quotient-metric} is a coordinate on the
asymptotic target geometry rather than a construction of the full cigar QFT.

\section{Abelian GLSM Coulomb-Branch One-Loop Calculation}
\label{sec:abelian-glsm-coulomb-one-loop-calculation}

The GLSM has a second protected local calculation beyond the classical Higgs
quotient.  Move to a Coulomb-branch region where
\(\sigma\neq0\), so charged chiral multiplets have masses
\[
  m_i(\sigma)=Q_i\sigma .
\]
The calculation below is a Wilsonian twisted \(F\)-term calculation whose
scope is the protected Coulomb coordinate chart.  The full infrared QFT
requires a separate continuum construction.

\begin{hypothesis}[Abelian Coulomb one-loop determinant]
\label{hyp:abelian-glsm-coulomb-one-loop-determinant}
Fix an abelian GLSM presentation as in
Definition~\ref{def:abelian-2d-glsm-presentation}, assume all charges \(Q_i\) are
nonzero, and restrict to a simply connected branch of the region
\[
  Q_i\sigma\neq0,\qquad i=1,\ldots,m.
\]
Assume a supersymmetric Wilsonian regulator preserving the twisted chiral
field strength \(\Sigma\), the gauge symmetry, and the integer charge
normalization.  The sign of the determinant logarithm is part of the component
linear-response calculation below; logarithm monodromy and compact theta
periodicity test that sign after it is derived but do not determine it.  In
the component convention used here, the finite part of the massive chiral
determinant is assumed to be represented by the local twisted superpotential
\begin{equation}
  \widetilde W_{\rm eff}(\Sigma)
  =
  -T\Sigma
  +
  \sum_{i=1}^m
  Q_i\Sigma\left(
    \log\frac{Q_i\Sigma}{\mu}-1
  \right)
  \label{eq:abelian-glsm-one-loop-twisted-superpotential}
\end{equation}
up to an additive constant and an affine function of \(\Sigma\).  The
renormalization scale \(\mu>0\) and the logarithm branches are part of the
scheme choice.  Boundary contributions from \(\sigma=0\), vortex sectors,
and singular loci are not included in this local determinant hypothesis.
\end{hypothesis}

\paragraph{Local determinant derivation.}
The part of Hypothesis~\ref{hyp:abelian-glsm-coulomb-one-loop-determinant}
that is a perturbative determinant calculation can be made explicit.  Work in
a background in which the twisted chiral field strength is slowly varying and
has lowest component \(\sigma\).  For a chiral multiplet of charge \(Q\), the
Coulomb-region complex mass is
\[
  M=Q\sigma .
\]
The local Wilsonian determinant is obtained by expanding the massive
Gaussian path integral to first order in the top component of \(\Sigma\), or
equivalently to first order in the background
\[
  \mathcal F_\Sigma=D+\ii F_{12}
\]
that is paired with the logarithmic FI-theta coordinate in the same component
normalization as Remark~\ref{rem:glsm-compact-fi-theta-normalization}.  At
\(\mathcal F_\Sigma=0\) the bosonic and fermionic determinants cancel.  The
gamma matrices and orientation are fixed by
\[
  \gamma^1=
  \begin{pmatrix}0&1\\ 1&0\end{pmatrix},
  \qquad
  \gamma^2=
  \begin{pmatrix}0&-\ii\\ \ii&0\end{pmatrix},
  \qquad
  \gamma_*=-\ii\gamma^1\gamma^2,
  \qquad
  \gamma^{12}=\frac12[\gamma^1,\gamma^2]=\ii\gamma_* .
\]
With this convention the terms linear in a constant charge-\(Q\) Euclidean
background follow from the quadratic charged-chiral action
\[
  S_{\Phi}^{(2)}
  =
  \int d^2x\,
  \left[
    \overline\phi
    \bigl(-\nabla_\mu\nabla_\mu+|M|^2+QD\bigr)\phi
    +
    \overline\psi\,\mathcal D_F\psi
    +
    \overline{\mathcal F}_\phi{\mathcal F}_\phi
  \right],
  \qquad
  \nabla_\mu=\partial_\mu+\ii Q A_\mu ,
\]
with
\[
  P_\pm=\frac12(1\pm\gamma_*),
  \qquad
  \mathcal D_F=\gamma^\mu\nabla_\mu+M P_+ +\overline M P_- .
\]
For constant \(M\) and slowly varying gauge field,
\([\nabla_1,\nabla_2]=\ii QF_{12}\), and the relevant second-order operators
are
\[
  \begin{aligned}
  \Delta_B
  &=
  -\nabla_\mu\nabla_\mu+|M|^2+QD,
  \\
  \mathcal D_F^\dagger\mathcal D_F
  &=
  -\nabla_\mu\nabla_\mu+|M|^2
  -\ii Q\gamma^{12}F_{12}
  +O(\nabla\sigma,D^2,F^2)
  \\
  &=
  -\nabla_\mu\nabla_\mu+|M|^2
  +Q\gamma_*F_{12}
  +O(\nabla\sigma,D^2,F^2).
  \end{aligned}
\]
where the omitted terms are either higher order in the slowly varying
background or cancel inside the supersymmetric supertrace.  The curvature
term is Hermitian.  The expression \(Q\gamma^{12}F_{12}\) without the factor
\(-\ii\) would be anti-Hermitian for real \(F_{12}\) and is not the positive
operator used here.  This formula is a component convention statement:
reversing the top-component convention of \(\Sigma\), or the chirality
convention in the massive fermion determinant, reverses the final holomorphic
logarithm and must be accompanied by the corresponding change in all Coulomb
equations.

The disputed real sign comes from differentiating the complex-scalar
determinant, not from branch periodicity.  Let \(K_\Lambda\) be any finite
gauge-covariant spectral cutoff for the scalar Laplacian, with eigenvalues
\(\lambda_n\).  A complex scalar is two real scalar fields, so its determinant
power is \(-\frac12-\frac12=-1\).  With the reference subtraction at \(\mu\),
\[
  \log Z_B^{\rm ren}(D)
  =
  -\sum_{\lambda_n\le\Lambda}
  \left[
  \log(\lambda_n+|M|^2+QD)
  -
  \log(\lambda_n+\mu^2+QD)
  \right],
\]
and therefore
\[
  \left.
  \frac{\partial\log Z_B^{\rm ren}}{\partial D}
  \right|_{D=0}
  =
  -Q\sum_{\lambda_n\le\Lambda}
  \left[
  \frac1{\lambda_n+|M|^2}
  -
  \frac1{\lambda_n+\mu^2}
  \right].
\]
Per unit area, the heat-kernel limit is the two-dimensional momentum integral
\begin{equation}
  \left.
  \frac1{{\rm Vol}}\,
  \frac{\partial\log Z_B^{\rm ren}}{\partial D}
  \right|_{D=0}
  =
  -Q\int\frac{d^2p}{(2\pi)^2}
  \left[
    \frac1{p^2+|M|^2}
    -
    \frac1{p^2+\mu^2}
  \right]
  =
  \frac{Q}{4\pi}\log\frac{|M|^2}{\mu^2}.
  \label{eq:abelian-glsm-determinant-local-density}
\end{equation}
Equivalently,
\begin{equation}
  \int_0^\infty
  \frac{\dd s}{4\pi s}\,
  \bigl(e^{-s|M|^2}-e^{-s\mu^2}\bigr)
  =
  \frac1{4\pi}\log\frac{\mu^2}{|M|^2}.
  \label{eq:abelian-glsm-proper-time-log}
\end{equation}
The equality follows by differentiating both sides with respect to
\(|M|^2\) and fixing the constant at \(|M|=\mu\):
\[
  \frac{\partial}{\partial |M|^2}
  \int_0^\infty\frac{\dd s}{s}
  \bigl(e^{-s|M|^2}-e^{-s\mu^2}\bigr)
  =
  -\int_0^\infty \dd s\,e^{-s|M|^2}
  =
  -\frac1{|M|^2}.
\]
Now apply the component pairing
\eqref{eq:glsm-twisted-f-component-pairing}.  A one-loop determinant is a
contribution to \(\log Z\), so its \(D\)-linear Wilsonian coefficient is read
from the same normalized top-component projection with the determinant sign
already contained in \(\log Z_B^{\rm ren}\).  If
\[
  \Re\left(\frac{\partial\widetilde W_{\rm 1\text{-}loop}}
    {\partial\Sigma}\right)
  =
  Q\log\frac{|M|}{\mu},
\]
then the \(D\)-density supplied by
\eqref{eq:glsm-twisted-f-component-pairing} is
\begin{equation}
  \frac1{2\pi}Q\log\frac{|M|}{\mu}
  =
  \frac{Q}{4\pi}\log\frac{|M|^2}{\mu^2},
  \label{eq:glsm-determinant-density-to-twisted-f}
\end{equation}
exactly the local density in
\eqref{eq:abelian-glsm-determinant-local-density}.  The factor \(1/(2\pi)\)
is therefore the Berezin/component normalization fixed by the FI and
\(-Q\Sigma Y\) terms, while the factor \(1/(4\pi)\) is the two-dimensional
momentum density.  Using \(1/(4\pi)\) again in the twisted-\(F\) projection
would miss a factor of two, and using \(\log|M|^2\) as the holomorphic
coefficient would double count the complex logarithm.  This extraction divides
by the signed analytic logarithm; a one-point sign test with a chosen ordering
of \(|M|\) and \(\mu\) would not determine the coefficient.

The \(F_{12}\) partner is a phase, not an independent real magnitude
calculation.  In the nonzero spectrum of \(\mathcal D_F^\dagger\mathcal D_F\),
states of opposite \(\gamma_*\)-chirality are paired, so the linear spin trace
of \(Q\gamma_*F_{12}\) does not change the determinant magnitude.  The
remaining phase is the Fujikawa Jacobian.  In a background with
\[
  k=\frac1{2\pi}\int F\in\mathbb Z,
\]
the same Hermitian square gives the local heat-kernel trace
\[
  \operatorname{Tr}_{\rm reg}\gamma_*
  =
  \lim_{s\downarrow0}
  \operatorname{Tr}\,
  \gamma_*\exp\{-s(-\nabla^2+Q\gamma_*F_{12})\}
  =
  -\frac{Q}{2\pi}\int F
  =
  -Qk .
\]
The companion finite regulator realizes the same sign with an \(L\times L\)
magnetic torus.  For flux \(Qk\) it uses link variables
\[
  U_x(x,y)=
  \exp\!\left(-\frac{2\pi\ii\,Qk\,y}{L^2}\right),
  \qquad
  U_y(x,y)=
  \begin{cases}
    1,& y<L-1,\\
    \exp\!\left(\frac{2\pi\ii\,Qk\,x}{L}\right),& y=L-1,
  \end{cases}
\]
so every plaquette has phase \(\exp(2\pi\ii Qk/L^2)\).  With the Hermitian
Wilson kernel \(H_W=\gamma_*(D_W-m)\), \(0<m<2\), the overlap index
\[
  \nu_{\rm ov}
  =
  -\frac12\operatorname{Tr}\operatorname{sign}(H_W)
\]
computes \(-Qk\) for the smooth finite-flux cells used here.  This finite
operator model fixes the chirality and flux convention independently of any
preassigned chiral dimension mismatch; the continuum index theorem is the
regulator-independent statement behind the heat-kernel limit.
The sign is fixed by the \(\gamma_*\) convention above: for \(Qk>0\), the
lowest Landau zero modes have negative \(\gamma_*\)-chirality.  If
\(M=|M|e^{\ii\alpha}\), the axial rotation that makes the mass real is
generated by \(-\gamma_*/2\), so the two half-angle measure contributions give
\[
  \exp\{-\ii\alpha\,\operatorname{Tr}_{\rm reg}\gamma_*\}
  =
  \exp\{\ii\alpha\,Qk\}.
\]
Since the Euclidean theta weight in
Remark~\ref{rem:glsm-compact-fi-theta-normalization} is
\(\exp(\ii\theta k)\), the same calculation shifts the compact theta
coordinate by
\[
  \theta\longmapsto\theta+Q\alpha .
\]
A one-sided measure, the opposite chirality convention, or an absolute-value
determinant would give respectively a factor of \(1/2\), the opposite sign, or
no theta shift.  Thus the component response and the Fujikawa Jacobian combine
into the holomorphic branch
\[
  Q\log\frac{M}{\mu}
  =
  Q\log\frac{Q\Sigma}{\mu}
\]
for the derivative of the twisted superpotential.  For \(m\) massive chiral
multiplets this gives, on the chosen simply connected Coulomb branch,
\begin{equation}
  \frac{\partial \widetilde W_{\rm eff}}{\partial\Sigma}
  =
  -T+\sum_{i=1}^m Q_i\log\frac{Q_i\Sigma}{\mu}.
  \label{eq:abelian-glsm-one-loop-derivative}
\end{equation}
Integrating \eqref{eq:abelian-glsm-one-loop-derivative} in the holomorphic
coordinate \(\Sigma\) gives
\[
  -T\Sigma+\sum_iQ_i\Sigma
  \left(\log\frac{Q_i\Sigma}{\mu}-1\right),
\]
up to an additive constant and an affine function of \(\Sigma\).  This is
\eqref{eq:abelian-glsm-one-loop-twisted-superpotential}.  The calculation is
local on the Coulomb chart: it uses \(Q_i\sigma\neq0\), chooses logarithm
branches, and says nothing by itself about vortex sectors, singular loci, or
the existence of the infrared continuum QFT.

\begin{controlledapproximation}[One-loop branch and monodromy ledger]
\label{ca:glsm-coulomb-one-loop-branch-monodromy}
Once the component determinant response has fixed the sign of the logarithm,
the branch ledger checks whether that sign is transported consistently by the
compact FI-theta coordinate.  It is not an independent sign derivation.  Work
on a simply connected Coulomb chart and choose branches
\[
  L_i(\Sigma)=\log_{\rm br}\frac{Q_i\Sigma}{\mu}.
\]
For a sign \(s_{\rm det}=\pm1\), the determinant contribution to the derivative
of the local twisted superpotential would be
\[
  \partial_\Sigma\widetilde W_{\rm 1\text{-}loop}
  =
  s_{\rm det}\sum_i Q_iL_i(\Sigma).
\]
The component calculation above selects \(s_{\rm det}=+1\) in this chapter's
convention.
A branch change
\[
  L_i\longmapsto L_i+2\pi\ii n_i,
  \qquad n_i\in\mathbb Z,
\]
changes the derivative by
\[
  s_{\rm det}\,2\pi\ii\sum_iQ_in_i .
\]
In the full Coulomb derivative
\(-T+s_{\rm det}\sum_iQ_iL_i\), this is exactly an integral-period shift of the
logarithmic FI coordinate,
\begin{equation}
  T\longmapsto T+s_{\rm det}\,2\pi\ii\sum_iQ_in_i .
  \label{eq:glsm-coulomb-branch-fi-shift}
\end{equation}
Since the physical fugacity is \(q=\ee^T\), the shift
\eqref{eq:glsm-coulomb-branch-fi-shift} leaves \(q\) invariant.  This is the
local form of the compact theta-period test.  A formula written with
\(\ee^\tau\), where \(\tau=\theta/(2\pi)+\ii r\), changes under the same
integer branch move and therefore lies outside the Coulomb fugacity convention
used here.

Going once around \(\Sigma=0\) changes every \(L_i\) by \(2\pi\ii\).  The
local determinant derivative therefore has monodromy
\[
  s_{\rm det}\,2\pi\ii\sum_iQ_i .
\]
This is the Coulomb-chart expression of the axial anomaly.  If
\(\sum_iQ_i=0\), the determinant derivative is single-valued after returning
to the same branch sheet.  If \(\sum_iQ_i\neq0\), the monodromy is compensated
by the same integral-period shift of \(T\); the exponentiated coordinate
\(\ee^T\) remains the compact gauge-flux fugacity.  Thus the branch data are
part of the regulated determinant coordinate before any dual variables
\(Y_i\) or primitive exponentials are introduced.

The primitive inherits the same ledger.  Under a branch change it shifts by
\[
  \Delta\widetilde W_{\rm eff}
  =
  s_{\rm det}\,2\pi\ii\,\Sigma\sum_iQ_in_i,
\]
an affine term in \(\Sigma\), and that affine term is precisely the local FI
coordinate shift in \eqref{eq:glsm-coulomb-branch-fi-shift}.  Notice the
important limitation: both signs give integral shifts of \(T\) and both leave
\(\ee^T\) single-valued.  Compact flux normalization fixes the period and the
\(2\pi\) factors; it does not by itself choose \(s_{\rm det}\).  The sign used
in \eqref{eq:abelian-glsm-one-loop-derivative} is therefore traced to the
component supertrace and mass-phase Jacobian above.  A calculation with
\(s_{\rm det}=-1\) would be a different determinant-response convention, not a
consequence of Hori--Vafa branch monodromy.
\end{controlledapproximation}

\paragraph{Coulomb critical equation.}
Under Hypothesis~\ref{hyp:abelian-glsm-coulomb-one-loop-determinant}, the
local Coulomb critical equation is
\begin{equation}
  -T+\sum_{i=1}^m Q_i\log\frac{Q_i\sigma}{\mu}=0 .
  \label{eq:abelian-glsm-coulomb-log-critical-equation}
\end{equation}
Equivalently, on the chosen logarithm branch,
\begin{equation}
  \prod_{i=1}^m
  \left(\frac{Q_i\sigma}{\mu}\right)^{Q_i}
  =
  \ee^T .
  \label{eq:abelian-glsm-coulomb-exponential-critical-equation}
\end{equation}
If all \(Q_i>0\), then this equation has
\begin{equation}
  Q_{\rm tot}=\sum_{i=1}^m Q_i
  \label{eq:abelian-glsm-positive-charge-vacuum-count}
\end{equation}
simple solutions for \(\sigma\) whenever \(\ee^T\neq0\).  For \(m=N\) fields
of charge \(1\), the count is \(N\).

Differentiate \eqref{eq:abelian-glsm-one-loop-twisted-superpotential} with
respect to \(\Sigma\).  For a single summand,
\[
  \frac{d}{d\Sigma}
  \left[
    Q_i\Sigma\left(\log\frac{Q_i\Sigma}{\mu}-1\right)
  \right]
  =
  Q_i\log\frac{Q_i\Sigma}{\mu}.
\]
The critical equation is therefore
\eqref{eq:abelian-glsm-coulomb-log-critical-equation}.  Exponentiating gives
\eqref{eq:abelian-glsm-coulomb-exponential-critical-equation}.

When all charges are positive, write
\[
  C_Q=\prod_{i=1}^m Q_i^{Q_i}.
\]
Then \eqref{eq:abelian-glsm-coulomb-exponential-critical-equation} becomes
\[
  C_Q\,\mu^{-Q_{\rm tot}}\sigma^{Q_{\rm tot}}=\ee^T .
\]
Since \(Q_{\rm tot}>0\) and the right side is nonzero, this polynomial
equation has \(Q_{\rm tot}\) nonzero roots.  Its derivative is proportional
to \(\sigma^{Q_{\rm tot}-1}\), so every nonzero root is simple.  For \(N\)
charge-one fields, \(Q_{\rm tot}=N\).

\begin{example}[\(\mathbb P^{N-1}\) GLSM quantum vacua]
\label{ex:cpn-glsm-coulomb-vacua}
Take \(m=N\), \(Q_i=1\), and \(W=0\).  The local equation is
\[
  \left(\frac{\sigma}{\mu}\right)^N=\ee^T .
\]
Thus the Coulomb-branch one-loop calculation predicts \(N\) massive twisted
chiral vacua in the chamber where this effective description is valid.  This
is the protected arithmetic behind the usual \(\mathbb P^{N-1}\) vacuum
count.  The statement that these vacua exhaust the continuum theory requires
the same regulator, compactness, and singular-locus controls stated in
Hypothesis~\ref{hyp:abelian-glsm-coulomb-one-loop-determinant}.
\end{example}

For the hypersurface charge vector \((1,\ldots,1,-d)\), the exponent of
\(\sigma\) in
\eqref{eq:abelian-glsm-coulomb-exponential-critical-equation} is \(N-d\),
the same integer as the axial anomaly coefficient
\eqref{eq:abelian-glsm-axial-anomaly}.  In the anomaly-free case \(N=d\),
the Coulomb equation becomes a condition on \(T\) and the logarithm branch
rather than an equation with finitely many isolated \(\sigma\)-solutions.
This is a protected signal of a singular locus in the FI-theta parameter
space.  A full infrared phase-transition statement requires control of the
continuum theory at the singular locus.

\paragraph{Classical chambers for the hypersurface GLSM.}
Fix integers \(N\geq2\) and \(d\geq2\).  Let the GLSM have chiral fields
\[
  X_1,\ldots,X_N,\quad P
\]
with charges
\begin{equation}
  Q(X_i)=1,\qquad Q(P)=-d,
  \label{eq:hypersurface-glsm-charges}
\end{equation}
and superpotential
\begin{equation}
  W=P\,G_d(X_1,\ldots,X_N),
  \label{eq:hypersurface-glsm-superpotential}
\end{equation}
where \(G_d\) is homogeneous of degree \(d\).  Assume that the projective
hypersurface
\begin{equation}
  Y_G=\{[X]\in\mathbb P^{N-1}:G_d(X)=0\}
  \label{eq:projective-hypersurface-phase-target}
\end{equation}
is smooth, equivalently the equations
\[
  G_d(X)=\partial_1G_d(X)=\cdots=\partial_NG_d(X)=0
\]
have no solution with \(X\neq0\).  Then the classical Higgs-vacuum quotient
has the following chamber description.
\begin{enumerate}
\item If \(r>0\), then \(X\neq0\), \(P=0\), and the quotient is
  \(Y_G\subset\mathbb P^{N-1}\).
\item If \(r<0\), then \(P\neq0\), \(X=0\) on the vacuum locus, and the
  unbroken finite gauge group is
  \[
    \mu_d=\{\zeta\in U(1):\zeta^d=1\}.
  \]
  Fluctuations of the \(X_i\) are therefore organized as a Landau--Ginzburg
  model with polynomial \(G_d(X)\), together with the residual finite
  \(\mu_d\) gauge quotient.
\item At \(r=0\), the point \(X=P=0\) obeys the \(D\)-term equation and the
  gauge quotient is singular.  A definition of the infrared QFT at the phase
  boundary requires additional continuum data beyond the chamber
  calculation.
\end{enumerate}

The \(F\)-term equations of
\eqref{eq:hypersurface-glsm-superpotential} are
\begin{equation}
  G_d(X)=0,\qquad P\,\partial_iG_d(X)=0,\quad i=1,\ldots,N.
  \label{eq:hypersurface-glsm-f-terms}
\end{equation}
The \(D\)-term equation is
\begin{equation}
  \sum_{i=1}^N|X_i|^2-d|P|^2=r.
  \label{eq:hypersurface-glsm-d-term}
\end{equation}

If \(r>0\), equation~\eqref{eq:hypersurface-glsm-d-term} forces
\(X\neq0\).  On the locus \(G_d(X)=0\), smoothness of \(Y_G\) implies that
the gradient \((\partial_iG_d(X))\) is not the zero vector.  Hence
\eqref{eq:hypersurface-glsm-f-terms} forces \(P=0\).  Quotienting nonzero
\(X\) by the charge-one \(U(1)\) action gives \([X]\in\mathbb P^{N-1}\), and
the remaining equation is \(G_d(X)=0\).

If \(r<0\), equation~\eqref{eq:hypersurface-glsm-d-term} forces \(P\neq0\).
Then \eqref{eq:hypersurface-glsm-f-terms} gives
\(\partial_iG_d(X)=0\) for all \(i\).  Homogeneity gives Euler's identity
\[
  d\,G_d(X)=\sum_{i=1}^N X_i\partial_iG_d(X),
\]
so \(G_d(X)=0\) follows as well.  By the smoothness assumption, the common
zero of \(G_d\) and all \(\partial_iG_d\) is \(X=0\).  A gauge
transformation \(\zeta\in U(1)\) preserves a nonzero value of \(P\) exactly
when \(\zeta^{-d}=1\), equivalently \(\zeta\in\mu_d\).  The transverse
coordinates are the \(X_i\), and the residual superpotential is \(G_d(X)\).

At \(r=0\), substituting \(X=0\) and \(P=0\) into
\eqref{eq:hypersurface-glsm-d-term} gives \(0=0\), and the \(F\)-term
equations are also satisfied.  The whole \(U(1)\) gauge group fixes this
point, so the quotient contains a positive-dimensional stabilizer stratum.
That stabilizer stratum is the singular locus of the classical quotient at
the phase boundary.

\paragraph{Adjunction, central charge, and the residual finite gauge theory.}
The smooth degree-\(d\) hypersurface
\[
  Y_G\subset\mathbb P^{N-1}
\]
has complex dimension \(N-2\).  Let
\[
  H=c_1(\mathcal O_{\mathbb P^{N-1}}(1)).
\]
The normal bundle of \(Y_G\) in \(\mathbb P^{N-1}\) is
\[
  N_{Y_G/\mathbb P^{N-1}}\cong \mathcal O(d)|_{Y_G}.
\]
The exact sequence
\[
  0\longrightarrow TY_G
  \longrightarrow T\mathbb P^{N-1}|_{Y_G}
  \longrightarrow \mathcal O(d)|_{Y_G}
  \longrightarrow0
\]
gives
\begin{equation}
  c_1(TY_G)=(N-d)H|_{Y_G},
  \qquad
  K_{Y_G}\cong \mathcal O(d-N)|_{Y_G}.
  \label{eq:hypersurface-adjunction-c1}
\end{equation}
Thus the topological Calabi--Yau condition in this hypersurface family is
\[
  d=N .
\]
In that case the large-volume sigma-model central-charge target is
\begin{equation}
  c_{\rm sigma}=3\dim_{\mathbb C}Y_G=3(N-2),
  \label{eq:hypersurface-sigma-central-charge-target}
\end{equation}
provided the infrared fixed point exists and realizes the standard
\(\mathcal N=(2,2)\) superconformal stress tensor.

On the Landau--Ginzburg side, the Fermat reference polynomial
\[
  G_d(X)=X_1^d+\cdots+X_N^d
\]
has quasihomogeneous weights \(q_i=1/d\), hence the protected
Landau--Ginzburg central-charge calculation gives
\begin{equation}
  c_{\rm LG}
  =
  3\sum_{i=1}^N\left(1-\frac{2}{d}\right)
  =
  3N\left(1-\frac{2}{d}\right).
  \label{eq:hypersurface-lg-central-charge}
\end{equation}
The equality
\[
  c_{\rm LG}=3(N-2)
\]
holds precisely when \(d=N\).  This is a protected numerical consistency
condition linking the two chambers.  The equality of central charges is one
piece of data in an infrared comparison; the equivalence of continuum QFTs
also requires the operator algebra, stress tensor, state space, and extended
sectors to be constructed and matched.

The finite group in the \(r<0\) chamber is an actual residual gauge group.
Choose a gauge in which the nonzero field \(P\) is fixed to a positive real
constant.  A gauge transformation \(\zeta\in U(1)\) preserves this gauge
choice exactly when
\[
  \zeta^{-d}=1.
\]
Therefore \(\zeta\in\mu_d\), and the remaining chiral fields transform as
\[
  X_i\longmapsto \zeta X_i .
\]
Homogeneity gives
\[
  G_d(\zeta X)=\zeta^dG_d(X)=G_d(X),
\]
so the Landau--Ginzburg superpotential descends to the finite gauge
quotient.  The finite-gauge QFT contains the invariant untwisted local
operators together with twisted sectors determined by the finite gauge
theory.  The invariant part of \(\operatorname{Jac}(G_d)\) is therefore only
one protected component of the orbifold theory; a full state-space
comparison must include the sectors created by the \(\mu_d\) defect lines.
For the Fermat Jacobi basis
\[
  X_1^{a_1}\cdots X_N^{a_N},
  \qquad 0\leq a_i\leq d-2,
\]
the untwisted invariant condition is
\[
  a_1+\cdots+a_N\equiv0\pmod d .
\]

\paragraph{Coulomb coordinate for the phase-boundary signal.}
For the charge vector \((1,\ldots,1,-d)\), the one-loop Coulomb equation
\eqref{eq:abelian-glsm-coulomb-exponential-critical-equation} becomes
\begin{equation}
  \left(\frac{\sigma}{\mu}\right)^N
  \left(\frac{-d\sigma}{\mu}\right)^{-d}_{\rm branch}
  =
  \ee^T .
  \label{eq:hypersurface-coulomb-charge-vector-equation}
\end{equation}
Equivalently,
\begin{equation}
  \left(\frac{\sigma}{\mu}\right)^{N-d}
  (-d)^{-d}_{\rm branch}
  =
  \ee^T .
  \label{eq:hypersurface-coulomb-power-equation}
\end{equation}
The subscript records the chosen logarithm branch for the negatively charged
field.  In the anomaly-free case \(d=N\), the \(\sigma\)-dependence cancels
and the local Coulomb chart selects the branch-dependent condition
\begin{equation}
  \ee^T=C_{\rm branch},
  \qquad
  C_{\rm branch}=(-N)^{-N}_{\rm branch}.
  \label{eq:hypersurface-coulomb-singular-coordinate}
\end{equation}
In the additive local FI coordinate
\[
  T_{\rm sing}=T-\log C_{\rm branch},
\]
the condition is \(T_{\rm sing}=0\).  At this value the one-loop large-\(\sigma\)
twisted superpotential has a critical direction rather than isolated local
Coulomb vacua.  This is the protected Coulomb-coordinate signal associated
with the chamber boundary.  The continuum behavior at that coordinate is a
separate infrared construction problem.

\begin{example}[The quintic GLSM as intrinsic two-dimensional QFT data]
\label{ex:quintic-glsm-intrinsic-qft}
Set \(N=d=5\).  The charges are
\[
  (1,1,1,1,1,-5),
\]
so the axial anomaly coefficient
\eqref{eq:abelian-glsm-axial-anomaly} is \(5-5=0\).  For \(r>0\), the
classical Higgs quotient is the degree-five hypersurface
\[
  G_5(X)=0
  \quad\text{inside}\quad
  \mathbb P^4.
\]
For \(r<0\), the residual finite gauge group is \(\mu_5\), and the local
Landau--Ginzburg polynomial is \(G_5(X_1,\ldots,X_5)\).  These are
statements about chambers of a two-dimensional supersymmetric gauge theory.
The associated infrared equivalence claim requires a two-dimensional
continuum construction and a comparison of its protected and unprotected
operator data.
\end{example}

\section{Infrared Comparison Problem}

For each two-dimensional supersymmetric model one must distinguish:
\[
\begin{gathered}
  \text{UV regulator},\qquad
  \text{supersymmetric Wilsonian flow},\\
  \text{protected operator algebra},\qquad
  \text{full infrared QFT}.
\end{gathered}
\]
Protected rings and indices can match before the full Hilbert space,
correlators, and stress tensor are identified.  A complete infrared
equivalence therefore requires a map of local algebras, states, stress
tensors, supercharges, and extended operators, with convergence or scaling
limits stated.

\begin{openproblem}[Two-dimensional supersymmetric RG construction]
\label{op:two-dimensional-susy-rg-construction}
Construct \(2D\) \(\mathcal N=(2,2)\) GLSM flows to Landau--Ginzburg and
hypersurface sigma-model fixed points with a regulator, supersymmetric
Wilsonian maps, protected-ring comparison, stress-tensor convergence, and
Hilbert-space reconstruction in one declared framework.  The target-space
terminology here refers only to an intrinsic two-dimensional QFT comparison
between continuum limits.
\end{openproblem}
