\chapter{Supersymmetric Gauge Theory}
\label{ch:susy-gauge-theory}

\ConventionsLorentzian

This chapter introduces supersymmetric gauge theory as field-variable data.
The particle spectrum, vacuum structure, Wilsonian effective action, and
nonperturbative dynamics are later constructions.  We keep the gauge
representative, the gauge-invariant observables, and the regularization
scheme separate.

\section{Gauge and Trace Conventions}
\label{sec:susy-gauge-trace-conventions}

This chapter uses the Hermitian-generator convention of
Chapter~\ref{ch:vol2-classical-yang-mills-matter}.  If
\(\operatorname{Lie}(G)\) is realized by anti-Hermitian matrices in a unitary
representation, write
\[
  \mathfrak g:=\ii\,\operatorname{Lie}(G)
\]
for the corresponding real vector space of Hermitian generators.  Its compact
Lie bracket in Hermitian coordinates is
\[
  [X,Y]_{\rm H}:=-\ii[X,Y],
  \qquad X,Y\in\mathfrak g,
\]
where the bracket on the right is the ordinary matrix commutator.  Fix an
invariant nondegenerate symmetric bilinear form
\(\operatorname{tr}:\mathfrak g\otimes\mathfrak g\to\R\).  In trace-delta
examples one chooses a basis \(t^a\) with
\(\operatorname{tr}(t^a t^b)=\delta^{ab}\).

In a local trivialization, a gauge field is a Hermitian
\(\mathfrak g\)-valued one-form \(A=A_\mu dx^\mu\), and the covariant
derivative on a field in a unitary representation is
\[
  \nabla_\mu=\partial_\mu-\ii A_\mu .
\]
The curvature is
\[
  F_{\mu\nu}
  :=
  \ii[\nabla_\mu,\nabla_\nu]
  =
  \partial_\mu A_\nu-\partial_\nu A_\mu-\ii[A_\mu,A_\nu].
\]
Under a smooth gauge transformation \(u:U\to G\),
\[
  A_\mu^u=uA_\mu u^{-1}+\ii u\,\partial_\mu u^{-1}.
\]

\begin{proposition}[Curvature covariance and gauge-invariant densities]
\label{prop:susy-gauge-curvature-covariance}
With the definitions above,
\[
  F_{\mu\nu}^u=uF_{\mu\nu}u^{-1}.
\]
Consequently
\(\operatorname{tr}(F_{\mu\nu}F^{\mu\nu})\) and
\(\operatorname{tr}(F_{\mu\nu}\widetilde F^{\mu\nu})\) are local
gauge-invariant scalar densities.
\end{proposition}

\begin{proof}
The transformation law for \(A_\mu\) is equivalent to
\[
  \nabla_\mu^u=u\nabla_\mu u^{-1}
\]
as an equality of differential operators.  Therefore
\[
  F_{\mu\nu}^u
  =
  \ii[\nabla_\mu^u,\nabla_\nu^u]
  =
  \ii u[\nabla_\mu,\nabla_\nu]u^{-1}
  =
  uF_{\mu\nu}u^{-1}.
\]
Invariance of the bilinear form gives
\(\operatorname{tr}(uXu^{-1}uYu^{-1})=\operatorname{tr}(XY)\).  Applying
this to \(X=F_{\mu\nu}\) and \(Y=F^{\mu\nu}\), and then to
\(Y=\widetilde F^{\mu\nu}\), proves the two density statements.
\end{proof}

The Yang--Mills coupling convention in this monograph is
\[
  \mathcal L_{\rm YM}
  =
  -\frac{1}{4g^2}\operatorname{tr}(F_{\mu\nu}F^{\mu\nu})
  +\frac{\theta}{32\pi^2}
  \operatorname{tr}(F_{\mu\nu}\widetilde F^{\mu\nu})
  +\cdots ,
\]
where \(\widetilde F^{\mu\nu}=\frac12
\epsilon^{\mu\nu\rho\sigma}F_{\rho\sigma}\) in the stated orientation
convention.  This convention applies uniformly to \(U(1)\), \(U(N)\), and
simple factors after the bilinear form has been fixed.

\section{Vector Multiplet as Field Variables}

\begin{definition}[Off-shell vector multiplet]
\label{def:off-shell-vector-multiplet}
An \(\mathcal N=1\) off-shell vector multiplet with gauge algebra
\(\mathfrak g\) consists locally of a connection \(A_\mu\), a Weyl spinor
\(\lambda_\alpha\) valued in \(\mathfrak g_\C\), its conjugate
\(\bar\lambda_{\dot\alpha}\), and a real auxiliary field \(D\) valued in
\(\mathfrak g\).  These are gauge-representative field variables.  The
physical local observables are gauge-invariant cohomological or operator
classes after quantization.
\end{definition}

In Wess--Zumino gauge, the supersymmetry transformations have the form
\[
  \delta A_\mu
  =
  \ii\epsilon\sigma_\mu\bar\lambda
  -\ii\lambda\sigma_\mu\bar\epsilon ,
\]
\[
  \delta\lambda_\alpha
  =
  (\sigma^{\mu\nu}\epsilon)_\alpha F_{\mu\nu}
  +\ii\epsilon_\alpha D,
  \qquad
  \delta D
  =
  -\epsilon\sigma^\mu D_\mu\bar\lambda
  -D_\mu\lambda\sigma^\mu\bar\epsilon .
\]
Here \(D_\mu X:=\partial_\mu X-\ii[A_\mu,X]\) for adjoint-valued component
fields \(X\).  The displayed transformations are not bare supertranslations
restricted to the Wess--Zumino slice.  They are bare supertranslations
followed by the compensating chiral gauge transformation that restores the
representative \eqref{eq:wess-zumino-gauge-vector-superfield}.

\begin{proposition}[Closure in Wess--Zumino gauge]
\label{prop:wess-zumino-gauge-closure}
Let \(\mathcal V\) be a local formal neighbourhood in the space of real
vector superfields, and let \(\mathcal S\subset\mathcal V\) be the
Wess--Zumino slice of
Proposition~\ref{prop:wess-zumino-gauge-representative}.  Assume the local
slice and the compensating chiral gauge parameters are chosen in this
neighbourhood.  For each supersymmetry parameter \(\epsilon\), write
\[
  \widehat\delta_\epsilon
  =
  \delta^0_\epsilon+\delta^{\rm g}_{\Lambda_\epsilon}
\]
for the vector field tangent to \(\mathcal S\), where
\(\delta^0_\epsilon\) is the supertranslation vector field on
\(\mathcal V\), and \(\delta^{\rm g}_{\Lambda_\epsilon}\) is the chiral
gauge vector field with parameter \(\Lambda_\epsilon(V)\).  Then, on
\(\mathcal S\),
\[
  [\widehat\delta_{\epsilon_1},\widehat\delta_{\epsilon_2}]
  =
  a^\rho\partial_\rho+\delta^{\rm g}_{\Omega_{12}},
  \qquad
  a^\rho=
  2\ii\left(
  \epsilon_1\sigma^\rho\bar\epsilon_2
  -\epsilon_2\sigma^\rho\bar\epsilon_1
  \right),
  \label{eq:wess-zumino-gauge-closure}
\]
where \(\Omega_{12}\) is a real residual ordinary gauge parameter determined
by the two compensators and the slice condition.  Equivalently, if
\(\Xi_{12}:=\Omega_{12}+a^\rho A_\rho\), then for every adjoint-valued
component field \(X\in\{\lambda,\bar\lambda,D\}\),
\[
  [\widehat\delta_{\epsilon_1},\widehat\delta_{\epsilon_2}]X
  =
  a^\rho D_\rho X+\ii[\Xi_{12},X],
\]
and for the connection
\[
  [\widehat\delta_{\epsilon_1},\widehat\delta_{\epsilon_2}]A_\mu
  =
  a^\rho F_{\rho\mu}+D_\mu\Xi_{12}.
\]
No Euler--Lagrange equation is used; the closure is off shell because the
auxiliary field \(D\) is part of the coordinate system.
\end{proposition}

\begin{proof}
On the full real-vector-superfield space \(\mathcal V\), the
supertranslation vector fields obey the algebra derived in
Chapter~\ref{ch:susy-superspace-superfields-local-actions}:
\[
  [\delta^0_{\epsilon_1},\delta^0_{\epsilon_2}]
  =
  a^\rho\partial_\rho .
\]
Let \(\mathcal D_{\rm g}\) be the vertical distribution generated by chiral
gauge transformations.  This distribution is involutive because chiral gauge
transformations form a group, and it is preserved by supertranslations:
if \(\bar D_{\dot\alpha}\Lambda=0\), then the supertranslation of
\(\Lambda\) is again a chiral gauge parameter.  Field dependence of
\(\Lambda_\epsilon(V)\) does not change this conclusion; differentiating a
chiral gauge parameter along any tangent vector in field space still gives a
chiral gauge parameter.  Therefore every term in the commutator
\[
  [\delta^0_{\epsilon_1}+\delta^{\rm g}_{\Lambda_{\epsilon_1}},
    \delta^0_{\epsilon_2}+\delta^{\rm g}_{\Lambda_{\epsilon_2}}]
\]
except the bare supertranslation commutator is vertical.  Hence the
commutator of the projected vector fields equals \(a^\rho\partial_\rho\)
modulo a chiral gauge vector field.

Both \(\widehat\delta_{\epsilon_i}\) are tangent to the Wess--Zumino slice,
so their commutator is tangent to that slice.  A vertical vector field that
is tangent to the Wess--Zumino slice is precisely the residual ordinary
gauge transformation left after the components \(C,\chi,M\) of a general
real vector superfield have been removed.  This gives
\eqref{eq:wess-zumino-gauge-closure}.  In Hermitian-generator conventions an
infinitesimal residual gauge parameter \(\Omega\) acts by
\[
  \delta^{\rm g}_\Omega A_\mu=D_\mu\Omega,
  \qquad
  \delta^{\rm g}_\Omega X=\ii[\Omega,X]
\]
on adjoint component fields.  The alternative covariant form follows from
the identity
\[
  a^\rho\partial_\rho A_\mu+D_\mu\Omega
  =
  a^\rho F_{\rho\mu}+D_\mu(\Omega+a^\rho A_\rho),
\]
which uses
\(F_{\rho\mu}=\partial_\rho A_\mu-\partial_\mu A_\rho
-\ii[A_\rho,A_\mu]\) and
\(D_\mu Y=\partial_\mu Y-\ii[A_\mu,Y]\).  This proves the stated closure
without a component equation of motion.
\end{proof}

Thus Wess--Zumino gauge is not a gauge-invariant description of
supersymmetry; it is a convenient representative after ordinary gauge
freedom has been partly fixed.

\section{Superfield Strength}

Let \(V\) be a real vector superfield.  For matrix notation in a
representation of \(G\), a chiral gauge transformation acts by
\[
  e^V\longmapsto e^{-\bar\Lambda}e^V e^{-\Lambda},
  \qquad \bar D_{\dot\alpha}\Lambda=0 .
\]
The chiral field strength is
\[
  W_\alpha
  =
  -\frac14\bar D^2\!\left(e^{-V}D_\alpha e^V\right),
  \qquad
  \bar D_{\dot\alpha}W_\beta=0.
\]
It transforms covariantly,
\[
  W_\alpha\longmapsto e^\Lambda W_\alpha e^{-\Lambda}.
\]
Consequently \(\operatorname{tr}(W^\alpha W_\alpha)\) is a chiral
gauge-invariant local superfield.  With the trace convention fixed above, the
gauge kinetic term is normalized so that its component expansion contains the
Yang--Mills term
\[
  -\frac{1}{4g^2}\operatorname{tr}(F_{\mu\nu}F^{\mu\nu}).
\]
All later anomaly and beta-function formulae must refer back to this coupling
normalization.

\section{Matter Multiplets and the Moment Map}

Let \(R\) be a finite-dimensional unitary complex representation of \(G\).
A chiral matter multiplet is a chiral superfield \(\Phi\) valued in \(R\).
Gauge covariance of the kinetic term is expressed by
\[
  \int d^4\theta\, \bar\Phi e^V\Phi .
\]
In components this term contains the scalar kinetic term, fermion kinetic
term, gauge couplings, and the auxiliary coupling
\[
  \langle\mu_R(\phi,\bar\phi),D\rangle,
\]
where \(\mu_R\in\mathfrak g^*\) is the moment map determined by
\[
  \langle \mu_R(\phi,\bar\phi),X\rangle
  =
  \phi^\dagger X_R\phi,
  \qquad X\in\mathfrak g .
\]
Here \(X_R\) is the Hermitian generator acting in \(R\).  A
Fayet--Iliopoulos parameter is a linear functional
\(\zeta\in\mathfrak g^*\) satisfying
\[
  \zeta([X,Y]_{\rm H})=0,\qquad X,Y\in\mathfrak g,
\]
so it factors through the abelian quotient of \(\mathfrak g\).  In
particular there is no FI parameter for a semisimple gauge algebra.

\begin{proposition}[Auxiliary \(D\)-field elimination]
\label{prop:susy-d-term-auxiliary-elimination}
Let \(\eta=\mu_R(\phi,\bar\phi)-\zeta\), with \(\zeta=0\) if no FI parameter
is allowed.  Use the bilinear form to define
\(\eta^\sharp\in\mathfrak g\) by
\[
  \operatorname{tr}(\eta^\sharp X)=\langle\eta,X\rangle
  \quad\text{for every }X\in\mathfrak g .
\]
For fixed scalar fields, the auxiliary part of the Lorentzian Lagrangian is
\[
  \mathcal L_D
  =
  \frac{1}{2g^2}\operatorname{tr}(D^2)
  +\langle\eta,D\rangle .
\]
Eliminating \(D\) gives
\[
  D=-g^2\eta^\sharp,
  \qquad
  \mathcal L_D\big|_{\rm on\ shell}
  =
  -\frac{g^2}{2}\operatorname{tr}(\eta^\sharp\eta^\sharp).
\]
Thus the corresponding scalar potential contribution is
\[
  V_D=\frac{g^2}{2}\operatorname{tr}(\eta^\sharp\eta^\sharp)
\]
in this convention.
\end{proposition}

\begin{proof}
By the definition of \(\eta^\sharp\),
\[
  \mathcal L_D
  =
  \frac{1}{2g^2}\operatorname{tr}(D^2)
  +\operatorname{tr}(\eta^\sharp D).
\]
Completing the square gives
\[
  \mathcal L_D
  =
  \frac{1}{2g^2}
  \operatorname{tr}\!\left(D+g^2\eta^\sharp\right)^2
  -\frac{g^2}{2}\operatorname{tr}(\eta^\sharp\eta^\sharp).
\]
The algebraic Euler--Lagrange equation is therefore
\(D+g^2\eta^\sharp=0\).  Substitution gives the displayed on-shell
Lagrangian.  Since Lorentzian scalar Lagrangians have the form kinetic minus
potential, the potential is the positive final expression.
\end{proof}

\section{\texorpdfstring{\(\mathcal N=2\) Gauge-Matter Coupling in
\(\mathcal N=1\) Variables}{N=2 Gauge-Matter Coupling in N=1 Variables}}

We shall often use \(\mathcal N=1\) superfields to describe
four-dimensional \(\mathcal N=2\) gauge theory.  The translation is
field-theoretic and does not use any string-theoretic construction.  Its
data are as follows.  Let \(G\) be compact, let \(R\) be a finite-dimensional
unitary representation, and choose the invariant trace form on
\(\mathfrak g\) used in the Yang--Mills kinetic term.  An
\(\mathcal N=2\) vector multiplet consists, in \(\mathcal N=1\) language, of
the vector superfield \(V\) together with an adjoint chiral superfield
\[
  \Phi=\Phi^a T^a\in\mathfrak g_{\mathbb C}.
\]
A hypermultiplet in \(R\) is represented by two chiral superfields
\[
  Q\in R,
  \qquad
  \widetilde Q\in R^\vee ,
\]
where \(R^\vee\) is the algebraic dual representation.  If
\(T_R^a\in\operatorname{End}(R)\) are the representation matrices, write
\[
  \Phi_R=\Phi^aT_R^a\in\operatorname{End}(R).
\]

With canonical flat chiral-field metrics and the Yang--Mills coupling
normalization already fixed in the vector kinetic term, the minimal
\(\mathcal N=2\) gauge-matter superpotential is
\begin{equation}
  W_{\mathcal N=2}
  =
  \sqrt2\,g_{\rm YM}\,\widetilde Q\,\Phi_R Q
  =
  \sqrt2\,g_{\rm YM}\,\Phi^a\,\widetilde Q\,T_R^a Q .
  \label{eq:n2-qcd-superpotential-n1-variables}
\end{equation}
The coefficient in \eqref{eq:n2-qcd-superpotential-n1-variables} is part of
the normalization statement: rescaling \(Q,\widetilde Q,\Phi\), or changing
the trace convention, changes the displayed coefficient together with the
kinetic terms and the gauge coupling.  Additional flavor tensors or
independent cubic coefficients are deformations of the \(\mathcal N=1\)
theory; they are not the minimal \(\mathcal N=2\) gauge coupling.

\begin{proposition}[Gauge invariance and \(F\)-term equations]
\label{prop:n2-qcd-superpotential-fterms}
Assume the preceding flat chiral-field metrics, and choose an orthonormal
basis \(T^a\) for the trace form.  Under a finite gauge transformation
\[
  Q\mapsto u_R Q,
  \qquad
  \widetilde Q\mapsto \widetilde Q u_R^{-1},
  \qquad
  \Phi_R\mapsto u_R\Phi_Ru_R^{-1},
\]
the contraction \(\widetilde Q\Phi_R Q\) is invariant.  The corresponding
scalar \(F\)-term potential is
\[
  V_F
  =
  2g_{\rm YM}^2
  \left(
    \|\varphi_R q\|_R^2
    +\|\widetilde q\,\varphi_R\|_{R^\vee}^2
    +\sum_a|\widetilde q\,T_R^a q|^2
  \right),
\]
where \(q,\widetilde q,\varphi\) are the scalar components of
\(Q,\widetilde Q,\Phi\), and
\(\varphi_R=\varphi^aT_R^a\).  Hence the \(F\)-flat equations are
\[
  \varphi_R q=0,
  \qquad
  \widetilde q\,\varphi_R=0,
  \qquad
  \widetilde q\,T_R^a q=0
  \quad\text{for every }a .
\]
\end{proposition}

\begin{proof}
Gauge invariance is the finite-dimensional identity
\[
  (\widetilde Q u_R^{-1})
  (u_R\Phi_Ru_R^{-1})
  (u_R Q)
  =
  \widetilde Q\Phi_RQ .
\]
Let \(c=\sqrt2\,g_{\rm YM}\).  In coordinates,
\[
  W=c\,\widetilde q_i(\varphi_R)^i{}_j q^j
    =c\,\varphi^a\,\widetilde q_i(T_R^a)^i{}_j q^j .
\]
Differentiating with respect to the three independent chiral coordinates
gives
\[
  \frac{\partial W}{\partial \widetilde q_i}
    =c(\varphi_R q)^i,\qquad
  \frac{\partial W}{\partial q^j}
    =c(\widetilde q\,\varphi_R)_j,\qquad
  \frac{\partial W}{\partial\varphi^a}
    =c\,\widetilde q\,T_R^a q .
\]
For canonical flat metrics the \(F\)-term potential is the squared norm of
this holomorphic gradient.  Since \(|c|^2=2g_{\rm YM}^2\), the displayed
potential follows, and its zero locus is precisely the simultaneous
vanishing of the three displayed gradient components.
\end{proof}

If \(R\) is the adjoint representation, the two chiral fields in the
hypermultiplet are also adjoint-valued.  Together with \(\Phi\) they are the
three adjoint chiral superfields of \(\mathcal N=4\) Yang--Mills written in
\(\mathcal N=1\) language, and the cubic term above becomes the usual
adjoint cubic after translating \(T_R^a\) into the adjoint action.  The
enhancement statement is a statement about the full kinetic, Yukawa, and
auxiliary-field normalizations together with the cubic polynomial.

\section{Gauge Anomalies as a Consistency Condition}

The classical field-variable construction is not yet a quantum gauge theory.
For a four-dimensional chiral gauge theory, gauge invariance of the quantum
theory requires the perturbative gauge anomaly class to vanish.  In a
representation \(R\), the cubic anomaly coefficient is the symmetric
trilinear form
\[
  A_R(X,Y,Z)
  =
  \operatorname{Tr}_R\!\left(X_R\{Y_R,Z_R\}\right),
  \qquad X,Y,Z\in\mathfrak g ,
\]
with the trace over chiral fermion variables.  The local anomaly cancellation
condition is
\[
  \sum_{\text{chiral fermions }i} A_{R_i}=0
\]
as an invariant symmetric trilinear form.  Global anomalies and mixed
anomalies require separate topological data and are not detected by this
polynomial alone.

\begin{proposition}[Conjugate-pair and real-representation cancellation]
\label{prop:susy-gauge-anomaly-conjugate-cancellation}
Let \(R^\vee\) be the conjugate representation of a unitary representation
\(R\), in the Hermitian-generator convention.  Then
\[
  A_{R^\vee}=-A_R .
\]
Consequently a vectorlike pair \(R\oplus R^\vee\) has zero perturbative local
gauge anomaly.  If \(R\) is equivalent to its conjugate representation, then
\(A_R=0\).  In particular, the adjoint gaugino in an \(\mathcal N=1\) vector
multiplet gives no perturbative cubic gauge anomaly.
\end{proposition}

\begin{proof}
If \(T_R^a\) are Hermitian generators in \(R\), then the conjugate
representation has Hermitian generators
\[
  T_{R^\vee}^a=-(T_R^a)^* .
\]
Therefore
\[
  \operatorname{Tr}_{R^\vee}
  \left(T_{R^\vee}^a\{T_{R^\vee}^b,T_{R^\vee}^c\}\right)
  =
  -
  \left[
  \operatorname{Tr}_{R}
  \left(T_R^a\{T_R^b,T_R^c\}\right)
  \right]^* .
\]
The trace on the right is real because the anticommutator of Hermitian
matrices is Hermitian and the invariant symmetric tensor is obtained by
symmetrizing over \(a,b,c\).  Hence \(A_{R^\vee}=-A_R\).  The vectorlike
statement follows by addition.  If \(R\simeq R^\vee\), anomaly coefficients
are representation invariants, so \(A_R=A_{R^\vee}=-A_R\), and \(A_R=0\).
The adjoint representation of a compact Lie algebra is real, giving the last
claim.
\end{proof}

\begin{openproblem}[Supersymmetric gauge theory with regulated Ward identities]
\label{op:susy-gauge-regulated-ward-identities}
Construct four-dimensional supersymmetric gauge theories with a regulator
that states the gauge symmetry, supersymmetry transformations, BV master
equation, local counterterm space, anomaly class, and decoupling limit at
finite cutoff.  This is the regulator-level input needed before exact
Wilsonian statements about holomorphy, moduli spaces, and duality can be
treated as quantum statements rather than coordinate calculations.
\end{openproblem}
