\chapter*{Volume Dependency Guide}
\addcontentsline{toc}{chapter}{Volume Dependency Guide}

This monograph is organized by logical dependency rather than by historical
order or course sequence.  A topic first appears where its primitive objects
are needed, but full development is placed where the required mathematical
machinery is available.  The list below is a reader-facing map for such
deferrals.

\begin{description}[leftmargin=2.7em,style=nextline]
\item[Scattering before perturbative amplitudes]
Particles, Haag--Ruelle wave operators, LSZ, and analytic scattering theory
are developed in Volume II.  Perturbative Feynman rules for S-matrix elements
are to be read only after that nonperturbative construction.

\item[Gauge theory and effective actions]
Gauge-fixed perturbation theory, BRST, BV, lattice gauge regulators,
anomalies, QCD, jets, energy correlators, and the hybrid Standard Model
framework are developed in Volume IV.  Wilsonian or 1PI claims for gauge
theories in Volume III are therefore conditional on either finite-cutoff BV
data or a lattice gauge-theory construction.

\item[Conformal fixed points]
The core CFT volume is Volume V.  Volume III discusses RG fixed points and
critical phenomena only to the extent needed for renormalization and scaling
limits.  Representation theory, radial quantization, OPE convergence,
crossing, conformal blocks, and light-ray observables belong to Volume V.

\item[Two-dimensional exact QFT]
Factorized scattering, Yang--Baxter consistency, form factors,
thermodynamic Bethe ansatz, exact finite-volume data, sine--Gordon,
Gross--Neveu, sigma-model families, and bridges to nonintegrable
two-dimensional theories are developed in Volume VI.

\item[Supersymmetry]
Supersymmetry algebras, particle versus field multiplets, superspace,
supersymmetric gauge theories, holomorphy, Wilsonian scheme questions,
Seiberg--Witten theory, localization caveats, and supersymmetric examples in
various dimensions are developed in Volume VII.

\item[Topological and cohomological theories]
Metric independence, bordism functoriality, extended TQFT, BF theory,
Chern--Simons theory, cohomological field theory, topological sigma models,
twists of supersymmetric theories, and Donaldson--Seiberg--Witten comparison
problems are developed in Volume VIII.

\item[Global structure and generalized symmetry]
Global forms of gauge groups, higher-form symmetries, line and surface
operators, confinement diagnostics, discrete theta terms, anomaly inflow,
defects, categorical symmetry, noninvertible defects, gauging, and SymTFT
viewpoints are developed in Volume IX.

\item[Thermal and hydrodynamic QFT]
KMS states, finite-temperature path integrals, Schwinger--Keldysh theory,
spectral functions, Kubo formulae, hydrodynamic constitutive relations,
thermal gauge theory, kinetic theory, anomalous transport, nonequilibrium
states, and long-time tails are developed in Volume X.

\item[Constructive, lattice, and numerical routes]
Constructive scalar models, lattice reflection positivity, continuum scaling
windows, Wilson lattice gauge theory, Monte Carlo, sign problems, rigorous RG,
lattice-to-continuum local QFT, stochastic quantization, Hamiltonian
truncation, DLCQ, and lattice fermion/chiral-symmetry questions are developed
in Volume XI.

\item[Curved backgrounds]
Locally covariant QFT, Hadamard states, point-splitting stress tensors, trace
anomalies, Unruh and Hawking effects, index-theoretic anomalies,
eta-invariants, cosmological spacetimes, microlocal spectrum conditions,
perturbative algebraic QFT on curved backgrounds, and semiclassical
backreaction are developed in Volume XII.
\end{description}

When a chapter uses a result from a later volume, the local statement should
identify the object being imported and its theorem status.  This guide only
locates the development; it does not prove a claim by itself.

\section*{Source Assembly Map}
\addcontentsline{toc}{section}{Source Assembly Map}

The source directory names still contain traces of earlier drafting stages.
They are therefore not authoritative indicators of the printed order.  The
printed parts are assembled by the following files, in the order shown in
\texttt{main.tex}.
\begin{center}
\begin{tabular}{c@{\quad}l}
Part I & \path{volumes/volume_i/volume_i_current.tex}\\
Part II & \path{volumes/volume_ii/volume_ii_current.tex}\\
Part III & \path{volumes/volume_iii/volume_iii_current.tex}\\
Part IV & \path{volumes/volume_iv/volume_iv_current.tex}\\
Part V & \path{volumes/volume_v/volume_v_current.tex}\\
Part VI & \path{volumes/volume_vi/volume_vi_current.tex}\\
Part VII & \path{volumes/volume_vii/volume_vii_current.tex}\\
Part VIII & \path{volumes/volume_viii/volume_viii_current.tex}\\
Part IX & \path{volumes/volume_ix/volume_ix_current.tex}\\
Part X & \path{volumes/volume_x/volume_x_current.tex}\\
Part XI & \path{volumes/volume_xi/volume_xi_current.tex}\\
Part XII & \path{volumes/volume_xii/volume_xii_current.tex}
\end{tabular}
\end{center}
For chapters not listed below, the chapter source file lies in the same
source-volume directory as the printed part.  The table records every current
exception.  A row marked ``section input'' means that the source file is
inserted into the preceding printed chapter and does not increment the chapter
counter.

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\begin{longtable}{>{\raggedright\arraybackslash}p{0.14\textwidth}
                  >{\raggedright\arraybackslash}p{0.28\textwidth}
                  >{\raggedright\arraybackslash}p{0.09\textwidth}
                  >{\raggedright\arraybackslash}p{0.39\textwidth}}
\hline
\textbf{Printed location} & \textbf{Printed title or source unit} &
\textbf{Source part} & \textbf{Source file}\\
\hline
\endhead
Part I, Ch. 2 & Wightman Fields and Reconstruction & IV &
\path{volumes/volume_iv/chapter01_wightman_fields_and_reconstruction.tex}\\
Part I, Ch. 3 & Algebraic QFT: Local Nets and States & IV &
\path{volumes/volume_iv/chapter03_algebraic_qft_local_nets_and_states.tex}\\
Part I, Ch. 4 & Superselection Sectors and Locality Properties & IV &
\path{volumes/volume_iv/chapter04_superselection_sectors_and_locality_properties.tex}\\
Part I, Ch. 12 & Osterwalder--Schrader Reconstruction & IV &
\path{volumes/volume_iv/chapter02_osterwalder_schrader_reconstruction.tex}\\
Part I, Ch. 13 & Kontsevich--Segal Functorial Quantum Field Theory & IV &
\path{volumes/volume_iv/chapter06_kontsevich_segal_functorial_qft.tex}\\
Part II, Ch. 17 & Haag--Ruelle Scattering Theory & I &
\path{volumes/volume_i/chapter12_haag_ruelle_scattering_theory.tex}\\
Part II, Ch. 18 & LSZ Reduction & I &
\path{volumes/volume_i/chapter13_lsz_reduction_and_the_s_matrix.tex}\\
Part II, Ch. 19 & Cross Sections, Phase Space, and Unitarity & I &
\path{volumes/volume_i/chapter14_cross_sections_partial_waves_and_unitarity.tex}\\
Part II, Ch. 20 & Massive Particles and Spin & I &
\path{volumes/volume_i/chapter15_massive_particles_with_spin.tex}\\
Part II, Ch. 21 & Massless Particles, Helicity, and Gauge Redundancy & I &
\path{volumes/volume_i/chapter17_massless_particles_helicity_and_gauge_redundancy.tex}\\
Part II, Ch. 22 & Haag--Ruelle Theory and Mathematical Scattering & IV &
\path{volumes/volume_iv/chapter05_haag_ruelle_and_mathematical_scattering.tex}\\
Part III, Ch. 30 & Generating Functionals and the One-Particle-Irreducible Effective Action & II &
\path{volumes/volume_ii/chapter23_generating_functionals_and_the_one_particle_irreducible_effective_action.tex}\\
Part III, Ch. 31 & Renormalizability and Local Counterterms & II &
\path{volumes/volume_ii/chapter08_renormalizability_and_local_counterterms.tex}\\
Part III, Ch. 32 & Subdivergences and Forest Formulas & II &
\path{volumes/volume_ii/chapter09_subdivergences_and_bphz_subtractions.tex}\\
Part III, Ch. 33 & The 1PI Renormalization Group & II &
\path{volumes/volume_ii/chapter10_renormalization_group_and_running_couplings.tex}\\
Part III, Ch. 34 & Renormalized Operators and Minimal Subtraction & II &
\path{volumes/volume_ii/chapter12_renormalized_operators_and_minimal_subtraction.tex}\\
Part III, Ch. 35 & Stress Tensor Trace and Conformal Currents & II &
\path{volumes/volume_ii/chapter13_stress_tensor_trace_scale_invariance_and_conformal_currents.tex}\\
Part III, Ch. 36 & Wilsonian Effective Actions and Exact Cutoff Flow & II &
\path{volumes/volume_ii/chapter16_wilsonian_effective_field_theory.tex}\\
Part III, Ch. 37 & The Wilson-Fisher Fixed Point and Scaling Operators & II &
\path{volumes/volume_ii/chapter14_the_wilson_fisher_fixed_point_and_scaling_operators.tex}\\
Part III, Ch. 38 & The Statistical Ising Model and Universality & II &
\path{volumes/volume_ii/chapter15_the_statistical_ising_model_and_universality.tex}\\
Part III, Ch. 39 & Effective Field Theories Without Poincare-Invariant UV Completion & II &
\path{volumes/volume_ii/chapter15b_efts_without_poincare_uv_completion.tex}\\
Part IV, Ch. 40 & Spinor Fields, Fermionic Statistics, and Grassmann Path Integrals & I &
\path{volumes/volume_i/chapter16_spinor_fields_fermionic_statistics_and_grassmann_path_integrals.tex}\\
Part IV, Ch. 40 & Spinor and Gamma-Matrix Conventions (section input) & I &
\path{volumes/volume_i/chapter16a_spinor_conventions.tex}\\
Part IV, Ch. 41 & Maxwell Theory, Constraints, and Gauge Fixing & I &
\path{volumes/volume_i/chapter18_maxwell_theory_constraints_and_gauge_fixing.tex}\\
Part IV, Ch. 42 & Quantum Electrodynamics and External States & I &
\path{volumes/volume_i/chapter19_quantum_electrodynamics_and_external_states.tex}\\
Part IV, Ch. 43 & QED Renormalization and Electromagnetic Form Factors & I &
\path{volumes/volume_i/chapter20_qed_renormalization_and_electromagnetic_form_factors.tex}\\
Part IV, Ch. 44 & Infrared Divergences and Inclusive QED & II &
\path{volumes/volume_ii/chapter22_infrared_divergences_and_inclusive_qed.tex}\\
Part IV, Ch. 45 & Classical Yang-Mills Theory and Matter Fields & II &
\path{volumes/volume_ii/chapter17_yang_mills_theory_and_matter_fields.tex}\\
Part IV, Ch. 46 & Lattice Yang--Mills as a Nonperturbative Formulation & II &
\path{volumes/volume_ii/chapter17b_lattice_yang_mills_nonperturbative_definition.tex}\\
Part IV, Ch. 47 & Gauge Fixing, Ghosts, and BRST Cohomology & II &
\path{volumes/volume_ii/chapter18_gauge_fixing_ghosts_and_brst_cohomology.tex}\\
Part IV, Ch. 48 & The BV Master Formalism for Gauge Effective Actions & II &
\path{volumes/volume_ii/chapter24_bv_master_formalism_gauge_effective_actions.tex}\\
Part IV, Ch. 49 & QCD Renormalization, Asymptotic Freedom, and Deep Inelastic Scattering & II &
\path{volumes/volume_ii/chapter19_qcd_renormalization_asymptotic_freedom_and_dis.tex}\\
Part IV, Ch. 50 & Jets, Infrared-Safe Observables, and Hadronization & II &
\path{volumes/volume_ii/chapter19b_jets_ir_safe_observables_and_hadronization.tex}\\
Part IV, Ch. 51 & Chiral Axial Anomalies & II &
\path{volumes/volume_ii/chapter20_chiral_axial_anomalies.tex}\\
Part IV, Ch. 52 & Instantons as Physical Amplitudes & II &
\path{volumes/volume_ii/chapter20d_instantons_and_physical_amplitudes.tex}\\
Part IV, Ch. 53 & The Schwinger Model: Two-Dimensional Quantum Electrodynamics & II &
\path{volumes/volume_ii/chapter20b_schwinger_model_two_dimensional_qed.tex}\\
Part IV, Ch. 54 & Large-\(N\) Two-Dimensional QCD and the Light-Front Bound-State Equation & II &
\path{volumes/volume_ii/chapter20c_large_n_two_dimensional_qcd_light_front.tex}\\
Part IV, Ch. 55 & The Standard Model as a Hybrid Quantum Field Theory & II &
\path{volumes/volume_ii/chapter19c_standard_model_hybrid_definition.tex}\\
Part IV, Ch. 56 & Global Anomalies, Spontaneous Symmetry Breaking, and Pions & II &
\path{volumes/volume_ii/chapter21_global_anomalies_spontaneous_symmetry_breaking_and_pions.tex}\\
Part V, Ch. 57 & Fixed Points and Conformal Data & III &
\path{volumes/volume_iii/chapter01_fixed_points_and_conformal_data.tex}\\
Part V, Ch. 58 & Conformal Killing Vectors and the Conformal Group & III &
\path{volumes/volume_iii/chapter02_conformal_killing_vectors_and_the_conformal_group.tex}\\
Part V, Ch. 59 & Stress Tensor, Weyl Structure, and Improvement & III &
\path{volumes/volume_iii/chapter03_stress_tensor_weyl_structure_and_improvement.tex}\\
Part V, Ch. 60 & Radial Quantization and the State-Operator Correspondence & III &
\path{volumes/volume_iii/chapter04_radial_quantization_and_state_operator_correspondence.tex}\\
Part V, Ch. 61 & Conformal Charges and Ward Identities & III &
\path{volumes/volume_iii/chapter05_conformal_charges_and_ward_identities.tex}\\
Part V, Ch. 62 & Primary Operators and Finite Conformal Transformations & III &
\path{volumes/volume_iii/chapter06_primary_operators_and_finite_transformations.tex}\\
Part V, Ch. 63 & Unitarity Bounds and Short Multiplets & III &
\path{volumes/volume_iii/chapter07_unitarity_bounds_and_short_multiplets.tex}\\
Part V, Ch. 64 & Correlation Functions and Conformal Frames & III &
\path{volumes/volume_iii/chapter08_correlation_functions_and_conformal_frames.tex}\\
Part V, Ch. 65 & The Operator Product Expansion & III &
\path{volumes/volume_iii/chapter09_operator_product_expansion.tex}\\
Part V, Ch. 66 & Light-Ray Operators and Energy Correlators & III &
\path{volumes/volume_iii/chapter10_light_ray_operators_and_energy_correlators.tex}\\
\hline
\end{longtable}
\endgroup
