\chapter{Relation Between Lattice And Continuum Local QFT}
\label{ch:constructive-lattice-continuum-local-qft}

\ConventionsEuclidean

This chapter explains how finite-lattice data become continuum local QFT
data.  The passage requires scaling maps for observables, convergence of
correlation functions as distributions, reflection positivity, and
reconstruction or local-net comparison.

\section{Scaling Maps}

Let \(a>0\) be the lattice spacing and let \(\Lambda_a=a\mathbb Z^d\cap V\)
inside a finite region \(V\).  A lattice field \(\phi_a(x)\) becomes a
continuum distribution only after a scaling map is specified:
\[
  \Phi_a(f)
  =
  Z_a^{1/2}a^d\sum_{x\in\Lambda_a} f(x)\phi_a(x),
\]
where \(f\in C_c^\infty(V)\) and \(Z_a\) is the field normalization.  Composite
operators require a mixing matrix
\[
  \mathcal O_{A,a}^{\mathrm{ren}}
  =
  \sum_B Z_{AB}(a)\mathcal O_{B,a}.
\]
The operator basis and normalization are part of the continuum-limit datum.

A useful way to remove a common normalization ambiguity is to use cell
averages.  Let \(Q_a(x)\) be the half-open cube of side \(a\) centered at
\(x\in a\mathbb Z^d\), and define
\[
  (I_af)(x)=a^{-d}\int_{Q_a(x)} f(u)\,d^du .
  \label{eq:cell-average-map}
\]
The smeared lattice field is then
\[
  \Phi_a(f)
  =
  Z_a^{1/2}a^d\sum_{x\in a\mathbb Z^d}(I_af)(x)\phi_a(x).
  \label{eq:cell-average-smeared-field}
\]
This is exactly the pairing of the lattice field with the piecewise-constant
test function
\[
  f_a(u)=\sum_{x\in a\mathbb Z^d}(I_af)(x)\mathbf 1_{Q_a(x)}(u).
  \label{eq:piecewise-constant-test-function}
\]

\begin{proposition}[Cell averages approximate test functions]
\label{prop:cell-average-test-function-convergence}
Let \(K\subset\mathbb R^d\) be compact and let \(f,h\in C_c^1(\mathbb R^d)\)
have support contained in \(K\).  Then
\[
  \left|
  a^d\sum_{x\in a\mathbb Z^d}(I_af)(x)(I_ah)(x)
  -
  \int_{\mathbb R^d} f(u)h(u)\,d^du
  \right|
  \leq
  C_K a
  \bigl(\|f\|_{C^1}\|h\|_{C^0}+\|f\|_{C^0}\|h\|_{C^1}\bigr),
  \label{eq:cell-average-product-error}
\]
where \(C_K\) depends only on a fixed compact enlargement of \(K\).  In
particular \(f_a\to f\) in \(L^2_{\mathrm{loc}}\).
\end{proposition}

\begin{proof}
On each cell \(Q_a(x)\),
\[
  |(I_af)(x)-f(u)|\leq c_d a\|f\|_{C^1},
  \qquad
  |(I_ah)(x)-h(u)|\leq c_d a\|h\|_{C^1}
  \]
for all \(u\in Q_a(x)\), by the mean-value theorem and
\(\operatorname{diam}Q_a(x)=\sqrt d\,a\).  Therefore
\[
  |(I_af)(x)(I_ah)(x)-f(u)h(u)|
  \leq
  c'_d a
  \bigl(\|f\|_{C^1}\|h\|_{C^0}+\|f\|_{C^0}\|h\|_{C^1}\bigr)
  +
  c''_d a^2\|f\|_{C^1}\|h\|_{C^1}.
  \]
Integrating this estimate over the cells meeting the compact enlargement of
\(K\) gives \eqref{eq:cell-average-product-error}, after increasing \(C_K\).
\end{proof}

\begin{definition}[Scaling-limit datum]
\label{def:scaling-limit-datum}
A scaling-limit datum for a lattice regulator family consists of:
\[
\begin{array}{ll}
\text{regulator spaces} & \Omega_a\text{ and probability states }\langle-\rangle_a,\\
\text{scaling trajectory} & \text{couplings and boundary conditions as
functions of }a,\\
\text{test spaces} & \mathcal D_i\text{ for each intended continuum field
or operator},\\
\text{operator maps} & \mathcal O_{i,a}(f)\text{ for }f\in\mathcal D_i,\\
\text{normalizations} & Z_{ij}(a)\text{ and operator-mixing matrices},\\
\text{topology} & \text{the seminorms in which all correlators converge}.
\end{array}
\]
Without these data the phrase ``the continuum limit of a lattice operator''
does not define a mathematical object.
\end{definition}

\section{Distributional Convergence}

The \(n\)-point Schwinger distributions are
\[
  S_{n,a}(f_1,\ldots,f_n)
  =
  \left\langle
  \Phi_a(f_1)\cdots \Phi_a(f_n)
  \right\rangle_a .
\]
A continuum Schwinger function \(S_n\) is obtained when
\[
  \lim_{a\to0}S_{n,a}(f_1,\ldots,f_n)
  =
  S_n(f_1,\ldots,f_n)
\]
for all test functions in the chosen class, with compatible infinite-volume
and continuum limits.  Pointwise lattice correlators are auxiliary data; the
continuum field is distributional.

The next criterion is the practical form in which distributional convergence
is often verified.  Let \(\mathcal D\) be a test-function space equipped with
an increasing family of seminorms \(p_1,p_2,\ldots\).  For functions
supported in a compact set \(K\), a typical seminorm is
\[
  p_{K,r}(f)=\max_{|\alpha|\leq r}\sup_{x\in K}|\partial^\alpha f(x)|.
\]

\begin{proposition}[Dense-test convergence plus uniform seminorm control]
\label{prop:dense-test-seminorm-convergence}
Let \(T_a\) be a family of \(n\)-linear functionals on \(\mathcal D\), and
let \(D_0\subset\mathcal D\) be dense in the seminorm used below.  Assume
there are a constant \(C\) and a continuous seminorm \(p\) such that
\[
  |T_a(f_1,\ldots,f_n)|
  \leq
  C\,p(f_1)\cdots p(f_n)
  \label{eq:uniform-multilinear-seminorm-bound}
\]
for all sufficiently small \(a\), and assume \(T_a(g_1,\ldots,g_n)\)
converges for all \(g_i\in D_0\).  Then there is a unique continuous
\(n\)-linear functional \(T\) satisfying
\[
  \lim_{a\to0}T_a(f_1,\ldots,f_n)=T(f_1,\ldots,f_n)
  \qquad(f_i\in\mathcal D).
  \label{eq:dense-test-extension-limit}
\]
\end{proposition}

\begin{proof}
For \(g_i\in D_0\), define \(T(g_1,\ldots,g_n)\) as the limit.  The bound
\eqref{eq:uniform-multilinear-seminorm-bound} passes to the limit on \(D_0\),
so \(T\) is uniformly continuous with respect to \(p\) on \(D_0^n\).  It
therefore has a unique continuous extension to \(\mathcal D^n\).  For
general \(f_i\in\mathcal D\), choose \(g_i\in D_0\) with
\[
  p(f_i-g_i)<\epsilon .
\]
Multilinearity and \eqref{eq:uniform-multilinear-seminorm-bound} give,
uniformly in \(a\),
\[
  |T_a(f_1,\ldots,f_n)-T_a(g_1,\ldots,g_n)|
  \leq
  C'\epsilon
  \prod_i(1+p(f_i))
  \]
after replacing one entry at a time.  The same estimate holds for \(T\).
Taking first \(a\to0\) on the dense tuple \(g_1,\ldots,g_n\), and then
\(\epsilon\to0\), proves \eqref{eq:dense-test-extension-limit}.
\end{proof}

This proposition isolates the countability step after the uniform seminorm
bound has been proved: convergence may be checked on a countable dense family
of test functions.  The hard constructive work is the bound
\eqref{eq:uniform-multilinear-seminorm-bound}, not the extension argument.

\section{Random-Walk Representation of a Massive Lattice Covariance}

The most concrete bridge from lattice locality to continuum estimates is the
path expansion of a massive lattice Green function.  Let \(G=(V,E)\) be a
finite undirected \(r\)-regular graph without self-loops, let \(A\) be its
adjacency matrix, and let \(m^2>0\).  Define
\[
  K_m=(m^2+r)I-A .
  \label{eq:massive-finite-graph-operator}
\]
For vertices \(x,y\), let \(N_n(x,y)\) be the number of length-\(n\) nearest
neighbor paths from \(x\) to \(y\), equivalently \(N_n(x,y)=(A^n)_{xy}\).

\begin{proposition}[Finite-graph random-walk resolvent]
\label{prop:finite-graph-random-walk-resolvent}
The matrix \(K_m\) is positive definite and
\[
  (K_m^{-1})_{xy}
  =
  \sum_{n=0}^{\infty}
  \frac{N_n(x,y)}{(m^2+r)^{n+1}} .
  \label{eq:finite-graph-random-walk-resolvent}
\]
Equivalently, if \(P=A/r\) is the simple-random-walk transition matrix and
\[
  \rho=\frac{r}{m^2+r}<1,
\]
then
\[
  (K_m^{-1})_{xy}
  =
  \frac1{m^2+r}\sum_{n=0}^{\infty}\rho^n P^n(x,y).
  \label{eq:finite-graph-random-walk-probability-form}
\]
\end{proposition}

\begin{proof}
For any vector \(u\),
\[
  u^*K_m u
  =
  m^2\sum_x |u_x|^2
  +
  \frac12\sum_{x\sim y}|u_x-u_y|^2 ,
  \label{eq:finite-graph-positive-quadratic-form}
\]
where each unoriented edge is counted once.  Thus \(K_m\) is positive
definite.  Since the operator norm of \(A\) on \(\ell^2(V)\) is at most \(r\),
\[
  \left\|\frac{A}{m^2+r}\right\|<1 .
\]
The Neumann series converges absolutely in operator norm:
\[
  K_m^{-1}
  =
  \frac1{m^2+r}
  \left(I-\frac{A}{m^2+r}\right)^{-1}
  =
  \sum_{n=0}^{\infty}\frac{A^n}{(m^2+r)^{n+1}} .
  \]
Taking the \((x,y)\) matrix entry gives
\eqref{eq:finite-graph-random-walk-resolvent}.  Replacing \(A^n\) by
\(r^nP^n\) gives \eqref{eq:finite-graph-random-walk-probability-form}.
\end{proof}

For a cubic lattice with spacing \(a\), the free covariance is obtained from
the same formula after the replacement
\[
  K_{m,a}=m^2I-a^{-2}\sum_{\mu=1}^{d}(T_\mu+T_\mu^{-1}-2I)
  =
  (m^2+2da^{-2})I-a^{-2}A .
  \label{eq:massive-cubic-lattice-operator}
\]
The path expansion then weights an \(n\)-step path by
\[
  a^{-2n}\bigl(m^2+2da^{-2}\bigr)^{-(n+1)}
\]
in the covariance matrix \(K_{m,a}^{-1}\).  In an interacting scalar
construction the free path expansion is replaced by a random-walk or polymer
expansion with local weights.  The proof obligations are the same in
structure as in Proposition~\ref{prop:finite-graph-random-walk-resolvent}:
identify a positive reference operator, expand the perturbation in local
objects, and prove a norm or polymer-sum bound that survives the intended
limit.

\section{Reflection Positivity And OS Reconstruction}

Assume the lattice measures are reflection positive with respect to a
hyperplane and that the scaling limit preserves reflection positivity:
\[
  \sum_{i,j}\overline c_i c_j\,S(F_i^\Theta F_j)\geq0
\]
for all finite linear combinations of positive-time test functionals.  If the
limit Schwinger functions also satisfy Euclidean covariance, symmetry,
regularity, and the OS growth hypotheses, OS reconstruction produces a
Hilbert space, vacuum, Hamiltonian, and Wightman distributions.  Thus the
lattice-to-continuum bridge is a theorem only after these hypotheses are
verified for the limiting distributions.

\begin{proposition}[Closedness of reflection positivity]
\label{prop:reflection-positivity-closed-under-limits}
Let \(S_a\) be reflection-positive Schwinger functionals on a common
positive-time test algebra \(\mathfrak A_+\).  Suppose that, for every
finite list \(F_1,\ldots,F_N\in\mathfrak A_+\), the limits
\[
  \lim_{a\to0} S_a(F_i^\Theta F_j)
  =
  S(F_i^\Theta F_j)
  \label{eq:reflection-positive-matrix-entry-limit}
\]
exist.  Then the limiting functional \(S\) is reflection positive:
\[
  \sum_{i,j=1}^{N}\overline c_i c_j\,S(F_i^\Theta F_j)\geq0
  \label{eq:limit-reflection-positivity}
\]
for all \(c_i\in\mathbb C\).
\end{proposition}

\begin{proof}
For each \(a\), the matrix
\[
  M_a(F)_{ij}=S_a(F_i^\Theta F_j)
\]
is positive semidefinite by reflection positivity.  Hence
\[
  \sum_{i,j}\overline c_i c_j M_a(F)_{ij}\geq0 .
\]
Taking the entrywise limit
\eqref{eq:reflection-positive-matrix-entry-limit} gives
\eqref{eq:limit-reflection-positivity}.  No compactness or reconstruction
input is hidden here; the only point is that positivity of finite Gram
matrices is a closed condition under entrywise convergence.
\end{proof}

\section{Local Algebras From Lattice Algebras}

For a bounded continuum region \(O\), let \(O_a\subset\Lambda_a\) be the set
of lattice sites or links whose cells lie in \(O\).  The lattice observable
algebra \(\mathcal A_a(O_a)\) is finite dimensional in spin systems and
finite-regulator gauge systems after gauge constraints are imposed.  A
continuum local algebra is approached by embeddings or correlation-function
limits
\[
  \mathcal A_a(O_a)\longrightarrow \mathcal A(O)
\]
that preserve inclusion, locality, and the state in the scaling limit.  The
construction is delicate for gauge theories because Gauss-law constraints and
center degrees of freedom affect the algebra assigned to a region with
boundary.

For spin systems the algebraic limit can be stated directly.  Let
\(\mathcal H_x\) be a finite-dimensional Hilbert space at each site and set
\[
  \mathcal A_a(X)=\bigotimes_{x\in X}\operatorname{End}(\mathcal H_x)
  \label{eq:finite-region-spin-algebra}
\]
for a finite set \(X\subset\Lambda_a\).  If \(X\subset Y\), the inclusion is
\[
  A\longmapsto A\otimes \mathbf 1_{Y\setminus X}.
\]
For disjoint \(X,Y\), the images of \(\mathcal A_a(X)\) and
\(\mathcal A_a(Y)\) commute.  The quasi-local algebra at fixed \(a\) is the
norm closure of the union over finite \(X\).

\begin{proposition}[Locality passes through state convergence for spin nets]
\label{prop:spin-locality-passes-to-gNS-limit}
Assume that for every bounded continuum region \(O\) and every sufficiently
small \(a\) there is a finite set \(O_a\subset\Lambda_a\), with isotony
\(O_1\subset O_2\Rightarrow O_{1,a}\subset O_{2,a}\).  Suppose a collection
of lattice observables \(A_{i,a}(O_i)\in\mathcal A_a(O_{i,a})\) has limiting
correlation functions in a state \(\omega\) on the abstract algebra generated
by the symbols \(A_i(O_i)\).  If \(O_1\) and \(O_2\) have positive
separation, then for all sufficiently small \(a\),
\[
  [A_{1,a}(O_1),A_{2,a}(O_2)]=0,
\]
and the limiting GNS representation obeys
\[
  [\pi_\omega(A_1(O_1)),\pi_\omega(A_2(O_2))]=0
  \]
on the dense domain generated by finite products of local symbols.
\end{proposition}

\begin{proof}
Positive separation of \(O_1\) and \(O_2\) implies
\(O_{1,a}\cap O_{2,a}=\varnothing\) for all sufficiently small \(a\), after
the cell approximation is chosen inside a fixed small thickening.  The finite
tensor-product construction then gives exact commutation at lattice spacing
\(a\).  For any finite product \(B_a\) of additional local lattice
observables whose correlations converge,
\[
  \omega\bigl(B[A_1(O_1),A_2(O_2)]\bigr)
  =
  \lim_{a\to0}\omega_a\bigl(B_a[A_{1,a}(O_1),A_{2,a}(O_2)]\bigr)
  =
  0 .
  \]
Moreover, for every such \(B\), set
\[
  C=[A_1(O_1),A_2(O_2)],
  \qquad
  C_a=[A_{1,a}(O_1),A_{2,a}(O_2)] .
\]
Then
\[
  \left\|\pi_\omega(C)\pi_\omega(B)\Omega_\omega\right\|^2
  =
  \omega(B^*C^*CB)
  =
  \lim_{a\to0}\omega_a(B_a^*C_a^*C_aB_a)
  =
  0 .
  \]
Thus the commutator annihilates the GNS dense domain generated by local
symbols.  Extension to bounded local von Neumann algebras, when they exist,
is a further closure statement and requires the chosen topology of the
scaling limit.
\end{proof}

Gauge theories are not obtained by the same tensor-factor argument after
imposing Gauss constraints.  The gauge-invariant algebra of a region need not
factor over a boundary; electric center, magnetic center, and extended
Hilbert-space prescriptions assign different boundary algebras.  A continuum
local-net claim for a lattice gauge scaling limit must therefore state which
regional algebra is being scaled and how charged or line-operator sectors are
represented.

\section{Gauge-Invariant Operators}

In lattice gauge theory, link variables are regulator coordinates.  Continuum
local gauge-invariant operators arise from small Wilson loops, covariant
lattice differences, and matter bilinears after subtracting and mixing all
operators with the same symmetries.  For example, the continuum action
density is represented by a plaquette combination
\[
  \mathcal P_a(x)
  =
  \frac{1}{a^4}\left(1-\frac{1}{\dim R}\operatorname{Re}\operatorname{Tr}_R
  U_{\square_x}\right)
  +\text{local mixings}.
\]
The mixings include identity and lower-dimensional terms allowed by the
regulator symmetries.

\section{Universality As A Convergence Statement}

Two lattice regulator families define the same continuum QFT when there are
scaling maps for their observables such that all reconstructed local QFT data
agree: Schwinger distributions, Wightman distributions or local nets,
symmetry actions, stress tensor normalization, line or defect sectors when
included, and background responses.  Equality of critical exponents is a
useful diagnostic but is not the full equivalence relation.

\section{Continuum Limit Checklist}

A lattice construction intended to define a continuum QFT must record:
\[
\begin{array}{ll}
\text{regulator family} & \text{lattice actions, measures, boundary
conditions},\\
\text{scaling trajectory} & \text{couplings as functions of }a,\\
\text{operator maps} & \text{normalization and mixing},\\
\text{limit topology} & \text{distributional, OS, Wightman, or local-net},\\
\text{positivity} & \text{reflection positivity or replacement structure},\\
\text{symmetries} & \text{continuum symmetry recovery and anomalies}.
\end{array}
\]

\begin{openproblem}[Local nets from interacting lattice gauge limits]
\label{op:local-nets-from-lattice-gauge-limits}
Construct continuum local algebras for interacting lattice gauge theory
scaling limits, including Gauss-law center choices, charged sectors,
extended operators, reflection positivity, and comparison with Wightman or OS
reconstruction.
\end{openproblem}
