\chapter{Lattice Reflection Positivity}
\label{ch:lattice-reflection-positivity}

\ConventionsEuclidean

Reflection positivity is the lattice analogue of the positivity axiom used in
Euclidean reconstruction.  It is a property of a regulated measure and a
chosen reflection, not an automatic property of every discretization.  This
chapter states the condition and verifies it for the basic nearest-neighbor
scalar lattice action.

\section{Reflection and Positive Algebra}

Let \(\Lambda\subset a\mathbb Z^D\) be a finite reflection-invariant lattice
with reflection \(\vartheta\) across the hyperplane \(x_0=0\).  Let
\(\phi_x\) be a real scalar variable at \(x\in\Lambda\).  Let
\(\mathcal A_+\) be the algebra of polynomial functions of the variables with
\(x_0>0\).  The reflection acts antilinearly on functions by
\[
  (\Theta F)(\phi)=\overline{F(\phi_{\vartheta x})}.
\]
For a probability measure \(d\mu(\phi)\), define
\[
  \langle F\rangle=\int F(\phi)\,d\mu(\phi).
\]

\begin{definition}[Lattice reflection positivity]
\label{def:lattice-reflection-positivity}
The measure \(d\mu\) is reflection positive with respect to
\(\vartheta\) if
\[
  \langle \Theta F\,F\rangle\ge 0
\]
for every \(F\in\mathcal A_+\).
\end{definition}

\section{Nearest-Neighbor Scalar Measure}

Consider the finite-volume lattice measure
\[
  d\mu(\phi)=Z^{-1}e^{-S(\phi)}\prod_{x\in\Lambda}d\phi_x
\]
with
\[
  S(\phi)=
  \frac12\sum_{\langle x y\rangle}c_{xy}(\phi_x-\phi_y)^2
  +\sum_x V(\phi_x),
  \qquad c_{xy}>0,
\]
where \(V\) is a real potential for which the finite-volume integral exists.
Split the action as
\[
  S=S_+ + \Theta S_+ + S_0
\]
where \(S_+\) contains terms strictly in the positive half, \(\Theta S_+\)
its reflected copy, and \(S_0\) the terms on or crossing the reflection
plane.  The crossing terms have the form
\[
  \frac12 c(\phi_x-\phi_{\vartheta x})^2
  =
  \frac12c\phi_x^2+\frac12c\phi_{\vartheta x}^2
  -c\phi_x\phi_{\vartheta x}.
\]
The first two terms can be assigned to \(S_+\) and \(\Theta S_+\).  The
coupling across the plane contributes the positive kernel
\[
  \exp(c\phi_x\phi_{\vartheta x})
  =
  \sum_{n=0}^{\infty}\frac{c^n}{n!}\phi_x^n\phi_{\vartheta x}^n .
\]

\begin{proposition}[Scalar lattice reflection positivity]
\label{prop:scalar-lattice-reflection-positivity}
The nearest-neighbor scalar lattice measure above is reflection positive.
\end{proposition}

\begin{proof}
After assigning all non-crossing and square terms to the two halves, the only
factor coupling reflected variables is a product of kernels
\(\exp(c_i\phi_i\phi_{\vartheta i})\) with \(c_i>0\).  Expanding each such
kernel in its power series gives a sum of terms
\[
  a_{\mathbf n}\,
  M_{\mathbf n}(\phi_+)\,
  \overline{M_{\mathbf n}(\phi_+)}
  \]
after reflection, with coefficients \(a_{\mathbf n}\ge 0\).  For
\(F\in\mathcal A_+\), the integral \(\langle \Theta F\,F\rangle\) becomes a
sum of absolute squares of half-lattice integrals weighted by
\(a_{\mathbf n}\).  Every term is nonnegative, hence the sum is nonnegative.
\end{proof}

\section{Free Lattice Fermions}

Reflection positivity for Grassmann variables is a positivity property of the
Berezin functional together with an antilinear anti-automorphism.  Let
\(\psi_A(x)\) and \(\bar\psi^A(x)\) be independent Grassmann-odd variables on
\(\Lambda\subset\mathbb Z^D\), and let \(C_\mu\) be Euclidean Clifford
matrices,
\[
  \{C_\mu,C_\nu\}_A{}^B=2\delta_{\mu\nu}\delta_A{}^B .
\]
The nearest-neighbor free lattice action is
\[
  -S
  =
  \sum_x\left[
  m\,\bar\psi^A(x)\psi_A(x)
  -\frac12\sum_{\mu=1}^D
  \bar\psi^A(x)(C_\mu)_A{}^B
  \bigl(\psi_B(x-e_\mu)-\psi_B(x+e_\mu)\bigr)
  \right].
\]
The corresponding two-point function is
\[
  \langle \bar\psi^B(y)\psi_A(x)\rangle
  =
  \int_{[-\pi,\pi]^D}\frac{d^Dp}{(2\pi)^D}\,
  e^{\ii p\cdot(x-y)}
  \frac{
  m\delta_A{}^B-\ii\sum_\mu(C_\mu)_A{}^B\sin p_\mu}
  {m^2+\sum_\nu\sin^2p_\nu}.
\]
Adding a Wilson term changes the numerator by a chirality-matrix term and
the denominator by the corresponding Wilson square; it removes doublers at
the price of explicitly coupling the reflection-positivity question to the
chosen Wilson operator and mass range.

For the naive action above, the reflection compatible with the simple
positivity proof is the reflection through the plane halfway between two
time slices.  Take \(v=e_1\), \(c=1/2\), and
\[
  R(x)=x-2(x\cdot e_1-c)e_1 .
\]
Let \(H=\{x:x_1\geq1\}\).  Define the antilinear anti-automorphism
\[
  \Theta(AB)=\Theta(B)\Theta(A),
\]
\[
  \Theta(\psi_A(x))=(C_1)_B{}^A\,\bar\psi^B(Rx),
  \qquad
  \Theta(\bar\psi^A(x))=(C_1)_A{}^B\,\psi_B(Rx).
\]
Reflection positivity means
\[
  \langle \Theta(\mathcal O)\mathcal O\rangle\geq0
\]
for every polynomial \(\mathcal O\) in the Grassmann variables supported in
\(H\), where positivity is the coefficient positivity obtained after the
Berezin integral is reduced to a sum of squares.

The placement of the reflection plane is essential.  A two-point test with
\(\mathcal O=\zeta^A\psi_A(x)\) gives, for the mid-link plane,
\[
  \int_{[-\pi,\pi]^D}\frac{d^Dp}{(2\pi)^D}\,
  \frac{|\zeta|^2\sin p_1\,\sin\!\bigl(p_1(2x_1-1)\bigr)}
  {m^2+\sum_\nu\sin^2p_\nu},
  \qquad x_1\geq1,
\]
which has the positive one-dimensional transfer-kernel expansion in the
normal direction.  For a site reflection \(c=0\), the same test contains the
term
\[
  \int_{[-\pi,\pi]^D}\frac{d^Dp}{(2\pi)^D}\,
  \frac{m\,(\zeta C_1\zeta^*)\cos(2p_1x_1)}
  {m^2+\sum_\nu\sin^2p_\nu},
\]
whose sign is not fixed for arbitrary spinor \(\zeta\).  A diagonal
reflection such as \(v=e_1+e_2\) similarly fails the same elementary
two-point positivity test.  Thus the fermion reflection structure is not
interchangeable among lattice hyperplanes.

\begin{proposition}[Mid-link free-fermion reflection positivity]
\label{prop:mid-link-free-fermion-reflection-positivity}
For the nearest-neighbor free lattice fermion action above, reflection
positivity holds for the mid-link reflection \(v=e_1,c=1/2\) with the
Grassmann reflection \(\Theta\) displayed above.
\end{proposition}

\begin{proof}
Write \(P_+=\{x:x_1=1\}\) and \(P_-=\{x:x_1=0\}\).  Split the action into
positive-half, reflected-half, and cross-plane terms.  The cross-plane
factor is a finite Grassmann polynomial over links crossing the plane:
\[
  \prod_{x\in P_+}
  \exp\!\left[
  -\frac12\bigl(
  \bar\psi^A(x)(C_1)_A{}^B\psi_B(Rx)
  -\bar\psi^A(Rx)(C_1)_A{}^B\psi_B(x)
  \bigr)\right].
\]
Using the displayed definition of \(\Theta\), each cross factor can be
written as a product of terms of the form
\[
  1+\frac12\left[
  \Theta(\psi_A(x))\,\bar\psi^A(x)
  +\Theta(\bar\psi^A(x))\,\psi_A(x)
  \right].
\]
Integrating the variables strictly inside \(H\) against
\(\mathcal O e^{-S_H}\) produces a boundary Grassmann polynomial
\(\Psi\) in the variables on \(P_+\).  The reflected half gives
\(\Theta(\Psi)\).  The remaining Berezin integral over \(P_+\cup P_-\) is a
finite sum of products \(\Theta(c_I)c_I\) of the coefficients of \(\Psi\),
with nonnegative numerical weights supplied by the cross factors.  This is
the Grassmann analogue of the scalar sum-of-squares argument.
\end{proof}

\section{Staggered Fermions and a Thirring Interaction}

In two dimensions with \(C_1=\sigma_3\) and \(C_2=\sigma_1\), a closed
staggered subsector of the Dirac lattice fermion can be written in terms of
one-component Grassmann variables \(\eta(x),\bar\eta(x)\).  In the convention
of the staggered reduction,
\[
  -S
  =
  \sum_{\langle xy\rangle}M(x,y)\bar\eta(x)\eta(y)
  +m\sum_x\bar\eta(x)\eta(x),
\]
with nearest-neighbor coefficients
\[
  M(x,x\pm e_1)=\pm\frac{(-1)^{x_2}}{2},
  \qquad
  M(x,x\pm e_2)=\pm1 .
\]
The mid-link reflection acts by
\[
  \Theta(\eta(x))=(-1)^{x_2}\bar\eta(Rx),
  \qquad
  \Theta(\bar\eta(x))=(-1)^{x_2}\eta(Rx).
\]
The same boundary-polynomial proof gives reflection positivity for the free
staggered action.

Add the nearest-neighbor Thirring interaction
\[
  -\Delta S
  =
  -U\sum_{x}\sum_{\mu=1,2}
  \bar\eta(x)\eta(x+e_\mu)\bar\eta(x+e_\mu)\eta(x).
\]
It is invariant under the displayed reflection.  The cross-plane factor has
the form
\[
  1+U\,\Theta(\eta(x)\bar\eta(x))\,\eta(x)\bar\eta(x)
\]
on each crossing link.  Therefore the same coefficient-square argument
extends to the interacting lattice Thirring model for
\[
  U\geq0 .
\]
This is a finite-regulator reflection-positivity statement.  Continuum
existence, scaling limits, and identification of the resulting QFT require
the additional compactness and renormalization estimates discussed in
Chapter~\ref{ch:constructive-continuum-limits-scaling-windows}.

\section{Transfer Matrix Interpretation}

When the lattice is periodic in Euclidean time, reflection positivity permits
the construction of a pre-Hilbert space from \(\mathcal A_+\) by quotienting
the null space of the sesquilinear form
\[
  (F,G)=\langle \Theta F\,G\rangle .
\]
Time translation by one lattice spacing acts as a contraction on this
pre-Hilbert space.  Under the additional hypotheses required for a transfer
matrix, this contraction is \(e^{-aH}\) for a positive self-adjoint lattice
Hamiltonian \(H\).  This is the finite-regulator version of the positivity
step in OS reconstruction.

\section{Gauge Fields}

For lattice gauge theory the configuration variables are group elements
\(U_\ell\in G\) on oriented links \(\ell\), with
\(U_{-\ell}=U_\ell^{-1}\).  Here \(G\) is a compact Lie group and the
finite-regulator measure is the product Haar measure multiplied by local
class functions of plaquette holonomies.  If
\[
  p=(\ell_1,\ell_2,\ell_3,\ell_4),
  \qquad
  U_p=U_{\ell_1}U_{\ell_2}U_{\ell_3}U_{\ell_4},
\]
then a one-plaquette gauge action is specified by a continuous positive
class function \(w:G\to\mathbb R_{>0}\):
\[
  d\mu(U)
  =
  Z^{-1}\prod_{\ell}dU_\ell\prod_p w(U_p).
\]
The reflection \(\vartheta\) maps links to reflected links with reflected
orientation.  The positive algebra \(\mathcal A_+\) is generated by matrix
elements of link variables supported in the positive half-lattice, and
\(\Theta\) is the antilinear anti-automorphism obtained by reflecting links
and complex conjugating coefficients.

The useful finite-volume theorem is not special to the defining
representation.  Let \(\widehat G\) be the set of equivalence classes of
finite-dimensional irreducible unitary representations of \(G\), let
\(\chi_\lambda\) be the character of \(\lambda\), and write
\(D^\lambda_{ij}\) for matrix elements in an orthonormal basis.

\begin{theorem}[Osterwalder-Seiler character criterion]
\label{thm:osterwalder-seiler-character-criterion}
Assume that the plaquette weight has an absolutely convergent character
expansion
\[
  w(g)=\sum_{\lambda\in\widehat G}c_\lambda\chi_\lambda(g),
  \qquad
  c_\lambda\geq0 .
\]
Then the finite-volume gauge measure defined above is reflection positive
with respect to the link-reflection plane.
\end{theorem}

\begin{proof}
Plaquettes lying strictly in the positive half-lattice and in the negative
half-lattice are assigned to \(W_+\) and \(\Theta W_+\), respectively.  A
plaquette crossing the reflection plane is cut into a positive half-path
\(c\) and its reflected half.  With a choice of orientation its holonomy has
the form
\[
  U_p=V_c\,\Theta(V_c)^{-1},
  \qquad
  V_c=\prod_{\ell\in c}U_\ell .
\]
Changing the orientation replaces \(g\) by \(g^{-1}\), which only conjugates
the character expansion and does not affect the positivity argument.

For one crossing plaquette, Peter-Weyl gives
\[
  w\!\left(V_c\Theta(V_c)^{-1}\right)
  =
  \sum_{\lambda}c_\lambda
  \sum_{i,j}
  D^\lambda_{ij}(V_c)\,
  \Theta\!\left(D^\lambda_{ij}(V_c)\right).
\]
The product over crossing plaquettes is therefore a finite or absolutely
convergent sum
\[
  \prod_{p\ {\rm crossing}}w(U_p)
  =
  \sum_\alpha C_\alpha\,\Theta(B_\alpha)B_\alpha,
  \qquad C_\alpha\geq0,
\]
where each \(B_\alpha\) is a polynomial in positive-half link matrix
elements.  For \(F\in\mathcal A_+\),
\[
  \langle \Theta F\,F\rangle
  =
  Z^{-1}\sum_\alpha C_\alpha
  \int
  \Theta(FW_+B_\alpha)\,(FW_+B_\alpha)
  \prod_{\ell}dU_\ell .
\]
Product Haar measure is invariant under reflection and factorizes between
the two halves once the crossing-plaquette expansion has been made.  Thus
each summand is the modulus square of the same half-lattice integral:
\[
  \int
  \Theta(FW_+B_\alpha)\,(FW_+B_\alpha)
  \prod_{\ell}dU_\ell
  =
  \left|
  \int_{\Lambda_+}FW_+B_\alpha\prod_{\ell\subset\Lambda_+}dU_\ell
  \right|^2
  \geq0 .
\]
Since \(C_\alpha\geq0\), the full sum is nonnegative.
\end{proof}

\begin{corollary}[Wilson plaquette action]
\label{cor:wilson-plaquette-reflection-positivity}
For \(G=SU(N)\) or \(U(N)\), the Wilson plaquette weight
\[
  w_W(g)
  =
  \exp\!\left(\frac{\beta}{N}\operatorname{Re}\operatorname{Tr}_F g\right),
  \qquad \beta\geq0,
\]
is reflection positive.
\end{corollary}

\begin{proof}
Let \(F\) be the defining representation and write
\(\chi_{\overline F}(g)=\overline{\chi_F(g)}=\chi_F(g^{-1})\).  Then
\[
  w_W(g)
  =
  \exp\!\left(\frac{\beta}{2N}
  \left(\chi_F(g)+\chi_{\overline F}(g)\right)\right)
  =
  \sum_{m,n\geq0}
  \frac{(\beta/2N)^{m+n}}{m!\,n!}\,
  \chi_F(g)^m\chi_{\overline F}(g)^n .
\]
The product \(\chi_F^m\chi_{\overline F}^n\) is the character of
\(F^{\otimes m}\otimes\overline F^{\otimes n}\).  Decomposing this finite
representation into irreducibles gives
\[
  \chi_F^m\chi_{\overline F}^n
  =
  \sum_{\lambda\in\widehat G}
  N^\lambda_{m,n}\chi_\lambda,
  \qquad
  N^\lambda_{m,n}\in\mathbb Z_{\geq0}.
\]
Hence the Wilson weight has nonnegative character coefficients.  Absolute
convergence follows from the uniform convergence of the exponential series
on compact \(G\), so Theorem~\ref{thm:osterwalder-seiler-character-criterion}
applies.
\end{proof}

For \(G=U(1)\), the same statement is the elementary Fourier expansion
\[
  e^{\beta\cos\theta}
  =
  \sum_{n\in\mathbb Z}I_n(\beta)e^{\ii n\theta},
  \qquad
  I_n(\beta)=\sum_{r\geq0}
  \frac{(\beta/2)^{2r+|n|}}{r!\,(r+|n|)!}\geq0 .
\]
This formula is the abelian form of the same positive-type condition.

The proof also explains why reflection positivity is not an automatic
consequence of gauge invariance.  Gauge fixing may destroy the product-Haar
factorization used above.  Improved gauge actions must be tested loop by
loop.  A sufficient condition is that every crossing-loop weight admit a
nonnegative character expansion after the loop is split into a positive
half-path and its reflected half.  The standard tree-level Symanzik or
Luscher-Weisz action and the Iwasaki action contain rectangle terms with
negative coefficient \(c_1\) in
\[
  S_{\rm imp}
  =
  -\frac{\beta}{N}
  \left(
  c_0\sum_{\rm plaquettes}\operatorname{Re}\operatorname{Tr}U_p
  +c_1\sum_{\rm rectangles}\operatorname{Re}\operatorname{Tr}U_R
  +\cdots
  \right),
\]
with \(c_1=-1/12\) at tree-level Symanzik improvement and
\(c_1\simeq -0.331\) for Iwasaki.  For a single crossing rectangle with
coefficient \(\alpha=\beta c_1/N<0\),
\[
  \exp\!\left(\alpha\operatorname{Re}\operatorname{Tr}_F g\right)
  =
  1+\frac{\alpha}{2}\left(\chi_F(g)+\chi_{\overline F}(g)\right)+O(\alpha^2),
\]
so the fundamental character coefficient is negative to first order.  Such
a factor is not of positive type and cannot enter the sum-of-squares proof.
Consequently the Wilson-action proof does not extend to these improved
one-step transfer matrices; any use of such actions in an argument requiring
reflection positivity must supply a separate positivity theorem, often after
blocking the time direction or changing the regulator.

\begin{openproblem}[Reflection-positive continuum limits]
\label{op:reflection-positive-continuum-limits}
Give verifiable criteria ensuring that a sequence of reflection-positive
lattice measures converges to OS-positive continuum Schwinger functions with
nontrivial local fields.  The criteria must include tightness or correlation
bounds, control of counterterms, and compatibility of reflection with the
continuum scaling limit.
\end{openproblem}
