\chapter{Rigorous Renormalization Group}
\label{ch:volume-xi-rigorous-renormalization-group}
\label{ch:constructive-rigorous-renormalization-group}

\ConventionsEuclidean

This chapter formulates renormalization group transformations as maps on
normed spaces of interactions or probability measures.  The aim is to replace
informal terminology about flows by explicit maps, estimates, fixed points,
and coordinate decompositions.

\section{Gaussian Block Integration}

Let \(C\) be a positive covariance on a finite regulator space of fields and
suppose it decomposes as
\[
  C=C_<+\Gamma .
\]
Let \(d\mu_C\), \(d\mu_{C_<}\), and \(d\mu_\Gamma\) be the corresponding
Gaussian measures.  For an interaction \(V\), define the single-step
effective interaction
\[
  e^{-V_<(\phi)}
  =
  \int d\mu_\Gamma(\zeta)\,e^{-V(\phi+\zeta)} .
\]
This identity is exact at finite regulator.  A scale transformation \(S_L\)
then gives an RG map
\[
  \mathcal R_L(V)=S_L V_< .
\]
The definition contains three inputs: covariance decomposition, integration
domain, and rescaling convention.

\section{Several RG Maps And Their Mathematical Objects}

The name ``renormalization group'' is used for several closely related, but
mathematically distinct, maps.  In this chapter the object being transformed
is always part of the statement.  Let \(C_\Lambda\) be a regulated covariance
and let \(C_{\Lambda_1}-C_{\Lambda_0}\) be positive for
\(\Lambda_1>\Lambda_0\) in the chosen cutoff ordering.  The exact Wilsonian
pushforward from \(\Lambda_1\) to \(\Lambda_0\) is the identity
\begin{equation}
  e^{-V_{\Lambda_0}(\phi)}
  =
  \int d\mu_{C_{\Lambda_1}-C_{\Lambda_0}}(\zeta)\,
  e^{-V_{\Lambda_1}(\phi+\zeta)} ,
  \label{eq:exact-wilsonian-pushforward-comparison}
\end{equation}
with additive normalization fixed separately.  Its semigroup property is the
Gaussian convolution identity:
\begin{equation}
  \int d\mu_{\Gamma_1}(\zeta_1)
  \int d\mu_{\Gamma_2}(\zeta_2)\,
  F(\phi+\zeta_1+\zeta_2)
  =
  \int d\mu_{\Gamma_1+\Gamma_2}(\zeta)\,
  F(\phi+\zeta),
  \label{eq:gaussian-convolution-semigroup-rg}
\end{equation}
valid whenever \(\Gamma_1\), \(\Gamma_2\), and
\(\Gamma_1+\Gamma_2\) are the corresponding positive covariances.  Thus the
exact Wilsonian map transforms interactions or measures.

The Wilson--Polchinski equation is the infinitesimal form of the same
Gaussian pushforward in a smooth-cutoff coordinate chart.  Let \(t\) be a
cutoff parameter such that lowering \(t\) by \(\delta t>0\) integrates the
positive shell covariance \(\delta t\,\dot C_t+O(\delta t^2)\), where
\(\dot C_t=\partial_t C_t\).  Thus
\begin{equation}
  e^{-V_{t-\delta t}[\phi]}
  =
  \left(1+\frac{\delta t}{2}\Delta_{\dot C_t}\right)
  e^{-V_t[\phi]}+O(\delta t^2).
  \label{eq:infinitesimal-gaussian-pushforward-taxonomy}
\end{equation}
Writing
\[
  \Delta_{\dot C_t}
  =
  \int_{x,y}\dot C_t(x,y)
  \frac{\delta}{\delta\phi(x)}
  \frac{\delta}{\delta\phi(y)} ,
\]
one obtains, equivalently, at finite regulator
\begin{equation}
  \partial_t e^{-V_t[\phi]}
  =
  -\frac12\,\Delta_{\dot C_t} e^{-V_t[\phi]} ,
  \label{eq:polchinski-exponential-form-rg-taxonomy}
\end{equation}
or, equivalently,
\begin{equation}
  \partial_t V_t
  =
  \frac12\langle \delta V_t,\dot C_t\,\delta V_t\rangle
  -\frac12\operatorname{Tr}(\dot C_t\,\delta^2 V_t),
  \label{eq:polchinski-action-form-rg-taxonomy}
\end{equation}
with this convention for \(t\).  Reparametrizing the scale variable multiplies
both sides by the corresponding chain-rule factor.  The equation is a
differential equation for the Wilsonian interaction.  Its theorem-level use
requires a function space, bounds on the functional derivatives, and a
construction of the trajectory.

The effective-average-action or Wetterich flow transforms a different
functional.  At finite regulator, introduce an infrared quadratic term
\[
  \Delta S_k[\phi]=\frac12\langle \phi,R_k\phi\rangle
\]
and define
\[
  W_k[J]=\log\int d\mu(\phi)\,
  \exp\!\left(\langle J,\phi\rangle-\Delta S_k[\phi]-V[\phi]\right).
\]
For \(\varphi=\delta W_k/\delta J\), the effective average action is the
modified Legendre transform
\begin{equation}
  \Gamma_k[\varphi]
  =
  \langle J,\varphi\rangle-W_k[J]-\Delta S_k[\varphi].
  \label{eq:effective-average-action-rg-taxonomy}
\end{equation}
Where the finite-dimensional Hessians are invertible, direct
differentiation gives
\begin{equation}
  \partial_t\Gamma_k[\varphi]
  =
  \frac12\operatorname{Tr}\!\left[
    \bigl(\Gamma_k^{(2)}[\varphi]+R_k\bigr)^{-1}\partial_tR_k
  \right].
  \label{eq:wetterich-rg-taxonomy}
\end{equation}
This is an exact identity for a regulated Legendre transform.  A fixed point
of a finite truncation of \(\Gamma_k\) is a computational approximation to
this functional flow; a theorem-level statement requires an embedding of the
truncation into a complete function space and a bound on the omitted
directions.

Constructive polymer RG transforms a pair consisting of local coordinates and
polymer activities.  Its single step has the form
\[
  (g,K)\longmapsto (g',K')
\]
where \(g'\) is obtained by a localization map extracting finitely many local
Taylor coefficients, while \(K'\) is the remaining nonlocal activity measured
in a polymer norm.  The decisive estimate is a contraction or summability
bound for \(K'\), together with Lipschitz dependence on \(g\).  This is the
form in which many rigorous scalar and fermionic constructions are proved.

Tensor RG starts from a tensor-network representation of a partition
function or state sum.  Its map acts on local tensors \(T\) rather than on a
continuum interaction:
\begin{equation}
  T\longmapsto \mathcal T(T),
  \label{eq:tensor-rg-map-taxonomy}
\end{equation}
with blocking, truncation or disentangling, and a declared Hilbert,
Hilbert--Schmidt, or Banach topology on the tensor space.  A tensor-RG fixed
point theorem therefore proves contraction or stable-manifold estimates for
\(\mathcal T\) in that tensor topology, plus a reconstruction map if the
claim concerns continuum QFT observables.

\begin{definition}[Comparison of two RG descriptions]
\label{def:comparison-of-rg-descriptions}
Let \((\mathcal B_1,\mathcal R_1)\) and
\((\mathcal B_2,\mathcal R_2)\) be two RG descriptions, each with its own
domain, norm, and reconstruction maps.  A comparison from the first to the
second consists of maps \(\Phi:\mathcal U_1\to\mathcal B_2\) and
\(\Psi:\mathcal O_1\to\mathcal O_2\), where \(\mathcal O_i\) are locally
convex spaces of observables with declared seminorm families
\(\mathfrak P_i\), together with estimates
\begin{equation}
  \|\Phi(\mathcal R_1 V)-\mathcal R_2(\Phi V)\|_{\mathcal B_2}
  \leq E_{\mathrm{RG}}(V),
  \label{eq:rg-description-intertwining-error}
\end{equation}
and
\begin{equation}
  p\!\left(
    \Psi\,\operatorname{Rec}^{(1)}_k(V),
    \operatorname{Rec}^{(2)}_k(\Phi V)
  \right)
  \leq E_{\mathrm{obs},p}(k,V),
  \label{eq:rg-description-observable-error}
\end{equation}
for every \(p\in\mathfrak P_2\), where \(E_{\mathrm{RG}}\) and
\(E_{\mathrm{obs},p}\) are explicitly bounded errors in the stated domains.
An exact equivalence is the special case in which both errors vanish and
\(\Phi\) identifies the relevant stable
manifold neighborhoods.  A controlled approximation is the case in which the
errors are shown to vanish in the continuum or long-distance limit relevant
to the claim.
\end{definition}

This definition is the standard used below when relating a perturbative
Polchinski trajectory, a BPHZ coordinate system, a functional-RG truncation,
or a tensor-network calculation to a Wilsonian fixed point.  The comparison
map and its error estimates are part of the mathematical claim.

\section{Banach Space Of Interactions}

Let \(\mathcal B\) be a Banach space of local functionals with norm
\(\|\cdot\|_{\mathcal B}\).  A typical constructive choice separates relevant
local coordinates from a polymer remainder:
\[
  V(\phi)=\sum_i g_i\mathcal O_i(\phi)+K(\phi),
\]
where \(g_i\in\R\) or \(\C\) and \(K\) is measured by a norm that includes
large-field weights and spatial decay.  The RG map is controlled on an open
set \(U\subset\mathcal B\) if
\[
  \mathcal R_L:U\longrightarrow\mathcal B
\]
is differentiable and its derivative satisfies explicit bounds in the chosen
norm.

\section{Linearization And Coordinates}

Let \(V_\ast\) be a fixed point:
\[
  \mathcal R_L(V_\ast)=V_\ast .
\]
The linearized map is
\[
  D\mathcal R_L|_{V_\ast}:\mathcal B\to\mathcal B .
\]
An eigenvector \(e_i\) with eigenvalue \(L^{y_i}\) is relevant, marginal, or
irrelevant according as \(y_i>0\), \(y_i=0\), or \(y_i<0\).  These words refer
to the declared linearized map and norm.  A coordinate \(g_i\) is a scaling
coordinate only after nonlinear terms have been put into a normal form or
controlled by a stable-manifold argument.

\begin{theorem}[Local stable graph for an RG map]
\label{thm:rg-local-stable-manifold}
Let \(\mathcal R\) be a \(C^1\) map on a Banach space \(\mathcal B\) with
fixed point \(V_\ast\).  Translate \(V_\ast\) to the origin and suppose
\[
  \mathcal B=E_u\oplus E_s
\]
where \(E_u\) is finite dimensional and \(E_s\) is closed.  In product norm
write
\begin{equation}
  \mathcal R(u,s)
  =
  (Au+F_u(u,s),\,Bs+F_s(u,s)),
  \label{eq:rg-stable-map-splitting}
\end{equation}
where \(A:E_u\to E_u\) is invertible, \(B:E_s\to E_s\), \(F(0)=0\),
and \(DF(0)=0\).  Assume that, on a closed ball of radius \(r\),
\begin{equation}
  \|A^{-1}\|\leq a<1,\qquad
  \|B\|\leq b<1,\qquad
  \|F(x)-F(y)\|\leq \kappa\|x-y\|.
  \label{eq:rg-stable-spectral-gap-constants}
\end{equation}
Choose \(\gamma\) with \(\max(a,b)<\gamma<1\) and suppose
\begin{equation}
  q_u:=\frac{\kappa a}{1-a\gamma}<1,
  \qquad
  q_s:=\frac{b+\kappa}{\gamma}<1,
  \qquad
  q:=\max(q_u,q_s)<1 .
  \label{eq:rg-stable-contraction-constants}
\end{equation}
Then, after possibly decreasing \(r\), for every sufficiently small
\(s_0\in E_s\) there is a unique forward orbit
\((u_n,s_n)_{n\geq0}\) satisfying \(s_0\) as its stable coordinate,
\[
  (u_{n+1},s_{n+1})=\mathcal R(u_n,s_n),
  \qquad
  \sup_{n\geq0}\gamma^{-n}\|(u_n,s_n)\|<\infty .
\]
The initial unstable coordinate \(u_0\) is a Lipschitz function \(h(s_0)\).
The graph \(u=h(s)\) is forward invariant, consists exactly of the points
whose forward orbit stays in the chosen \(\gamma\)-decay neighborhood, and
is tangent to \(E_s\) at the fixed point in the sense
\(\|h(s)\|/\|s\|\to0\) as \(s\to0\).
\end{theorem}

\begin{proof}
Fix \(\rho\leq r\) and \(\|s_0\|\leq \rho\).  Let
\[
  \mathcal X_{\gamma,\rho}(s_0)
  =
  \left\{
    x=(u_n,s_n)_{n\geq0}:
    s_0\ {\rm is\ fixed},\
    \|x\|_\gamma:=\sup_{n\geq0}\gamma^{-n}\|(u_n,s_n)\|\leq\rho
  \right\}.
\]
With the metric induced by \(\|\cdot\|_\gamma\), this is a closed subset of
the Banach space of \(\gamma\)-weighted sequences, hence is complete.  Define
the Lyapunov--Perron map \(\mathcal T_{s_0}\) by
\begin{align}
  (\mathcal T_{s_0}x)^s_0&=s_0,\\
  (\mathcal T_{s_0}x)^s_{n+1}
    &=B s_n+F_s(u_n,s_n), \label{eq:rg-stable-lp-s-part}\\
  (\mathcal T_{s_0}x)^u_n
    &=
    -\sum_{j=0}^{\infty}
      A^{-(j+1)}F_u(u_{n+j},s_{n+j}).
      \label{eq:rg-stable-lp-u-part}
\end{align}
The series in \eqref{eq:rg-stable-lp-u-part} converges absolutely because
\begin{equation}
  \gamma^{-n}
  \sum_{j\geq0}
  \|A^{-(j+1)}\|\,\|F_u(u_{n+j},s_{n+j})\|
  \leq
  \frac{\kappa a}{1-a\gamma}\|x\|_\gamma .
  \label{eq:rg-stable-u-series-bound}
\end{equation}
Here \(F(0)=0\) was used.  Similarly,
\begin{equation}
  \gamma^{-(n+1)}
  \|B s_n+F_s(u_n,s_n)\|
  \leq
  \frac{b+\kappa}{\gamma}\|x\|_\gamma .
  \label{eq:rg-stable-s-recursion-bound}
\end{equation}
Equations \eqref{eq:rg-stable-contraction-constants},
\eqref{eq:rg-stable-u-series-bound}, and
\eqref{eq:rg-stable-s-recursion-bound} show that
\(\mathcal T_{s_0}\) maps \(\mathcal X_{\gamma,\rho}(s_0)\) into itself
for every \(\rho\leq r\) with \(\|s_0\|\leq\rho\).

For two sequences \(x,y\in\mathcal X_{\gamma,\rho}(s_0)\), the same
estimates applied to differences give
\begin{equation}
  \|\mathcal T_{s_0}x-\mathcal T_{s_0}y\|_\gamma
  \leq
  q\|x-y\|_\gamma ,
  \label{eq:rg-stable-lp-contraction}
\end{equation}
with \(q<1\).  Banach's fixed-point theorem gives a unique fixed sequence
\(x(s_0)\).  The \(s\)-part of the fixed-point equation is exactly the
stable coordinate recurrence.  The \(u\)-part is equivalent to the unstable
coordinate recurrence because multiplying
\eqref{eq:rg-stable-lp-u-part} by \(A\) and shifting the summation gives
\[
  A u_n+F_u(u_n,s_n)=u_{n+1}.
\]
Thus the fixed sequence is precisely a forward orbit of \(\mathcal R\) with
\(\gamma\)-decay.  Conversely, any forward orbit with this decay must satisfy
\eqref{eq:rg-stable-lp-u-part}; otherwise the growing homogeneous solution
of \(u_{n+1}=Au_n+F_u(u_n,s_n)\) would violate
\(\sup_n\gamma^{-n}\|u_n\|<\infty\).

Define \(h(s_0)=u_0(s_0)\).  If \(s_0,t_0\) are two stable initial
coordinates, the fixed-point equations give
\[
  \|x(s_0)-x(t_0)\|_\gamma
  \leq
  \|s_0-t_0\|+q\|x(s_0)-x(t_0)\|_\gamma ,
\]
and hence
\begin{equation}
  \|x(s_0)-x(t_0)\|_\gamma
  \leq
  \frac{1}{1-q}\|s_0-t_0\|.
  \label{eq:rg-stable-graph-lipschitz-bound}
\end{equation}
Therefore \(h\) is Lipschitz.  Since \(DF(0)=0\), the Lipschitz constant
\(\kappa\) for \(F\) on a ball of radius \(\rho\) tends to zero as
\(\rho\downarrow0\).  The bound
\eqref{eq:rg-stable-u-series-bound} then implies
\(\|h(s)\|/\|s\|\to0\), so the graph is tangent to \(E_s\) at the fixed
point.  Forward invariance is built into the recurrence, and the
\(\gamma\)-weighted bound gives convergence to the fixed point.
\end{proof}

\begin{theorem}[\(C^1\) stable graph from the Lyapunov--Perron equation]
\label{thm:rg-c1-stable-graph-criterion}
Assume the hypotheses and notation of
Theorem~\ref{thm:rg-local-stable-manifold}.  Suppose in addition that
\(F\) is \(C^1\) on the ball used there and that \(DF\) is uniformly
continuous on that ball.  Let
\[
  \kappa_\rho
  =
  \sup_{\|(u,s)\|\leq \rho}\|DF(u,s)\|_{\mathcal B\to\mathcal B}.
\]
After restricting to a smaller ball if necessary, assume
\begin{equation}
  q_\rho
  =
  \max\!\left\{
    \frac{a\kappa_\rho}{1-a\gamma},
    \frac{b+\kappa_\rho}{\gamma}
  \right\}
  <1 .
  \label{eq:rg-c1-stable-graph-small-derivative}
\end{equation}
Then the orbit map
\[
  s_0\longmapsto x(s_0)=(u_n(s_0),s_n(s_0))_{n\geq0}
  \in\mathcal X_{\gamma,\rho}(s_0)
\]
constructed in Theorem~\ref{thm:rg-local-stable-manifold} is \(C^1\) as a
map from the stable-coordinate ball into the \(\gamma\)-weighted sequence
space.  Its derivative in the direction \(v\in E_s\) is the unique
\(\gamma\)-weighted sequence \(Y(v)=(Y^u_n(v),Y^s_n(v))_{n\geq0}\) satisfying
\begin{align}
  Y^s_0(v)&=v, \label{eq:rg-c1-stable-derivative-initial}\\
  Y^s_{n+1}(v)
  &=
  B\,Y^s_n(v)
  +
  DF_s(x_n(s_0))\,Y_n(v),                                      \\
  Y^u_n(v)
  &=
  -\sum_{j=0}^{\infty}
  A^{-(j+1)}DF_u(x_{n+j}(s_0))\,Y_{n+j}(v).
  \label{eq:rg-c1-stable-derivative-unstable}
\end{align}
Consequently the stable graph \(h(s_0)=u_0(s_0)\) is \(C^1\), with
\[
  Dh(s_0)v=Y^u_0(v),
  \qquad
  Dh(0)=0 .
\]
\end{theorem}

\begin{proof}
Let \(X_\gamma\) be the Banach space of sequences
\((u_n,s_n)_{n\geq0}\) with finite norm
\[
  \|x\|_\gamma
  =
  \sup_n\gamma^{-n}\|(u_n,s_n)\|.
\]
The
Lyapunov--Perron map used in the proof of
Theorem~\ref{thm:rg-local-stable-manifold} may be written
\[
  x=\mathcal T(s_0,x).
\]
For fixed \(s_0\), its derivative with respect to \(x\) is obtained by
replacing each occurrence of \(F\) in
\eqref{eq:rg-stable-lp-s-part}--\eqref{eq:rg-stable-lp-u-part} by
\(DF(x_n)\).  The same estimates as
\eqref{eq:rg-stable-u-series-bound} and
\eqref{eq:rg-stable-s-recursion-bound}, with \(\kappa\) replaced by
\(\kappa_\rho\), give
\begin{equation}
  \|D_x\mathcal T(s_0,x)\|_{X_\gamma\to X_\gamma}
  \leq q_\rho<1 .
  \label{eq:rg-c1-lp-linearized-contraction}
\end{equation}
The derivative of \(\mathcal T\) with respect to the parameter \(s_0\) is the
sequence whose stable component at \(n=0\) is the input vector and whose
other components are zero before the contraction equation is solved.

For \(v\in E_s\), define \(Y(v)\) to be the fixed point of the affine
linear equation
\begin{equation}
  Y
  =
  D_{s_0}\mathcal T(s_0,x(s_0))v
  +
  D_x\mathcal T(s_0,x(s_0))Y .
  \label{eq:rg-c1-affine-derivative-equation}
\end{equation}
Banach's fixed-point theorem applies because of
\eqref{eq:rg-c1-lp-linearized-contraction}.  Written in components,
\eqref{eq:rg-c1-affine-derivative-equation} is exactly
\eqref{eq:rg-c1-stable-derivative-initial}--\eqref{eq:rg-c1-stable-derivative-unstable}.
It also gives the bound
\[
  \|Y(v)\|_\gamma\leq\frac{\|v\|}{1-q_\rho}.
\]

It remains to check that \(Y(v)\) is the derivative of \(x(s_0)\).  Put
\[
  \Delta_t=\frac{x(s_0+tv)-x(s_0)}{t}.
\]
Subtract
\eqref{eq:rg-c1-affine-derivative-equation} from the difference quotient of
\[
  x(s_0+tv)=\mathcal T(s_0+tv,x(s_0+tv)),
  \qquad
  x(s_0)=\mathcal T(s_0,x(s_0)).
\]
The result is
\[
  \Delta_t-Y(v)
  =
  D_x\mathcal T(s_0,x(s_0))(\Delta_t-Y(v))
  +
  r_t,
\]
where \(r_t\to0\) in \(X_\gamma\).  The convergence of \(r_t\) is the
Fréchet differentiability of \(F\) on each finite component together with
the uniform continuity of \(DF\), the uniform \(\gamma\)-weighted bound on
the two orbits, and the dominated geometric sums in
\eqref{eq:rg-stable-u-series-bound}.  Taking norms and using
\eqref{eq:rg-c1-lp-linearized-contraction} gives
\[
  \|\Delta_t-Y(v)\|_\gamma
  \leq
  q_\rho\|\Delta_t-Y(v)\|_\gamma+\|r_t\|_\gamma ,
\]
so \(\Delta_t\to Y(v)\).  Thus \(x(\cdot)\) is differentiable.

Continuity of \(Dx(s_0)\) follows from the same contraction equation:
if \(s_0\) and \(s_0'\) are close, the corresponding orbits are close in
\(X_\gamma\), and uniform continuity of \(DF\) makes
\[
  D_x\mathcal T(s_0,x(s_0))
  -
  D_x\mathcal T(s_0',x(s_0'))
\]
small as an operator on \(X_\gamma\).  The inverse
\((1-D_x\mathcal T)^{-1}\) is uniformly bounded by
\((1-q_\rho)^{-1}\), so the fixed solution of
\eqref{eq:rg-c1-affine-derivative-equation} depends continuously on
\(s_0\).

Finally, at \(s_0=0\) the orbit is identically zero and \(DF(0)=0\).  The
derivative equations reduce to
\[
  Y^s_n=B^n v,\qquad Y^u_n=0 .
\]
Therefore \(Dh(0)v=Y^u_0(v)=0\) for every \(v\in E_s\).
\end{proof}

\section{Critical Surfaces And Continuum Tuning}

The local stable graph is the mathematical object behind continuum tuning.
Let the hypotheses and notation of
Theorem~\ref{thm:rg-local-stable-manifold} hold.  Write
\[
  \Pi_u:\mathcal B=E_u\oplus E_s\longrightarrow E_u,
  \qquad
  \Pi_s:\mathcal B=E_u\oplus E_s\longrightarrow E_s
\]
for the continuous projections.  In a sufficiently small neighborhood of
\(V_\ast\), define the local critical surface
\begin{equation}
  \mathcal W^s_{\rm loc}(V_\ast)
  =
  \left\{
    V_\ast+h(s)+s:\ s\in E_s,\ \|s\|\ {\rm small}
  \right\}.
  \label{eq:local-critical-surface-rg}
\end{equation}
The word ``critical'' here means attraction to the fixed point under the
declared RG map.  It is a property of the pair
\((\mathcal R,\mathcal B)\), together with the reconstruction maps if a
continuum QFT limit is claimed.

\begin{definition}[Tuning map]
\label{def:rg-tuning-map}
In the setting above, the local tuning map is
\begin{equation}
  \tau(V)
  =
  \Pi_u(V-V_\ast)-h\!\left(\Pi_s(V-V_\ast)\right),
  \label{eq:rg-tuning-map}
\end{equation}
defined for \(V\) in the local product neighborhood on which the graph
\(h\) exists.  A point \(V\) is tuned to the local critical surface precisely
when
\begin{equation}
  \tau(V)=0 .
  \label{eq:rg-tuning-equation}
\end{equation}
If \(\dim E_u=r\) and a basis of \(E_u\) has been chosen, this is an
\(r\)-component equation.  The components of \(\tau\), rather than a chosen
list of formal Lagrangian coefficients, are the intrinsic relevant tuning
coordinates in this RG chart.
\end{definition}

\paragraph{Where the analytic burden lies.}
Theorem~\ref{thm:rg-local-stable-manifold} proves the existence of a Lipschitz
stable graph under the displayed Lyapunov--Perron contraction hypotheses.
Theorem~\ref{thm:rg-c1-stable-graph-criterion} gives one self-contained
route to \(C^1\) regularity: prove uniform derivative bounds for the chosen
Banach RG map so that the differentiated Lyapunov--Perron equation remains a
contraction.  A model-specific RG theorem must establish these derivative
bounds in its polymer, kernel, tensor, lattice, or BV norm.  The geometric
consequence after this analytic input is a standard submersion calculation,
not an RG theorem in its own right.

\paragraph{Submanifold form after differentiability.}
Assume the stable graph \(h:E_s\to E_u\) in
Theorem~\ref{thm:rg-local-stable-manifold} is \(C^1\) on a neighborhood of
the origin in \(E_s\), and let \(r=\dim E_u<\infty\).  Then
\(\mathcal W^s_{\rm loc}(V_\ast)\) is a \(C^1\) submanifold of
\(\mathcal B\) of codimension \(r\).  Equivalently, the equation
\(\tau(V)=0\) cuts out the local critical surface by \(r\) independent
conditions.

Let \(P\) be a finite-dimensional \(C^1\) parameter manifold of microscopic
finite-regulator theories, and let
\[
  \iota:P\longrightarrow\mathcal B
\]
be a \(C^1\) chart map into the RG Banach space.  If \(p_0\in P\) satisfies
\(\tau(\iota(p_0))=0\) and
\begin{equation}
  D(\tau\circ\iota)_{p_0}:T_{p_0}P\longrightarrow E_u
  \label{eq:rg-tuning-transversality}
\end{equation}
has rank \(r\), then the tuned microscopic parameters near \(p_0\),
\begin{equation}
  P_{\rm tuned}
  =
  \{p\in P:\tau(\iota(p))=0\},
  \label{eq:tuned-microscopic-parameter-surface}
\end{equation}
form a \(C^1\) submanifold of \(P\) of codimension \(r\).

For \(V=V_\ast+u+s\), with \(u\in E_u\) and \(s\in E_s\), the derivative of
\(\tau\) in the \(E_u\)-direction is the identity map on \(E_u\):
\begin{equation}
  D_u\tau(V_\ast+u+s)=1_{E_u}.
  \label{eq:tuning-map-submersion}
\end{equation}
Thus \(\tau\) is a \(C^1\) submersion onto the finite-dimensional space
\(E_u\).  Its zero set is exactly
\[
  \{V_\ast+u+s:\ u=h(s)\},
\]
which is the graph \eqref{eq:local-critical-surface-rg}.  The submersion
theorem for Banach spaces with finite-dimensional target gives a \(C^1\)
submanifold of codimension \(\dim E_u=r\).  In this product chart the
submanifold statement is also direct: the map
\[
  s\longmapsto V_\ast+h(s)+s
\]
parametrizes the zero set, and the transverse coordinates are precisely
\(\tau\in E_u\).

For the microscopic parameter space, apply the finite-dimensional implicit
function theorem to \(\tau\circ\iota:P\to E_u\).  The rank condition
\eqref{eq:rg-tuning-transversality} makes \(0\in E_u\) a regular value near
\(p_0\).  Hence the inverse image
\((\tau\circ\iota)^{-1}(0)\) is a \(C^1\) submanifold of \(P\) of codimension
\(r\), which is \eqref{eq:tuned-microscopic-parameter-surface}.

\paragraph{Quantitative finite-dimensional tuning estimate.}
The rank condition is not the estimate used in a finite-regulator
construction.  After a choice of \(r\) microscopic tuning coordinates,
one needs a ball on which the nonlinear tuning equation is solved with
controlled constants.  Let \(X\) and \(Y\) be \(r\)-dimensional normed
spaces, let \(F:\overline B_\rho(0)\subset X\to Y\) be \(C^1\) on a
neighborhood of the closed ball, and let
\[
  e=F(0),\qquad A=DF(0).
\]
Assume \(A:X\to Y\) is invertible and, for some \(0\leq\kappa<1/2\),
\begin{equation}
  \|A^{-1}e\|\leq\frac{\rho}{2},
  \qquad
  \sup_{\|x\|\leq\rho}\|A^{-1}(DF(x)-A)\|\leq\kappa .
  \label{eq:quantitative-tuning-contraction-hypotheses}
\end{equation}
Then \(F(x)=0\) has a unique solution in the closed ball
\(\overline B_\rho(0)\).  Indeed define
\[
  T(x)=x-A^{-1}F(x).
\]
For \(\|x\|,\|y\|\leq\rho\), the fundamental theorem of calculus gives
\begin{align}
  T(x)-T(y)
  &=
  -\int_0^1
  A^{-1}\bigl(DF(y+t(x-y))-A\bigr)(x-y)\,\dd t ,
  \label{eq:tuning-contraction-difference}
\end{align}
so \(\|T(x)-T(y)\|\leq\kappa\|x-y\|\).  Also
\[
  \|T(x)\|
  \leq
  \|T(0)\|+\kappa\|x\|
  =
  \|A^{-1}e\|+\kappa\|x\|
  \leq
  \rho .
\]
Thus \(T\) is a contraction of the complete metric space
\(\overline B_\rho(0)\) into itself, and Banach's fixed-point theorem gives
a unique fixed point in that closed ball.  The fixed-point equation
\(T(x)=x\) is exactly \(F(x)=0\).

In the RG application \(F\) is the local coordinate expression for
\(\tau\circ\iota\) in the chosen \(r\) microscopic tuning variables, after
any spectator parameters have been fixed or placed in a small auxiliary
ball.  The residual \(\|A^{-1}e\|\) measures how far the proposed finite
microscopic point is from the local critical surface in intrinsic relevant
coordinates; the second bound in
\eqref{eq:quantitative-tuning-contraction-hypotheses} is the quantitative
transversality estimate.  This estimate still does not construct a fixed
point or a continuum QFT.  It only says that, once the stable graph and the
microscopic parameter chart have already been established with these
constants, the finite-dimensional relevant-coordinate equation can actually
be solved in the declared neighborhood.

\paragraph{Linear finite-scale tuning exponents.}
Let \(E_u\) be a finite-dimensional normed vector space with basis
\(\{e_a\}_{a=1}^r\).  Suppose the linearized unstable map is diagonal in this
basis,
\[
  A e_a=\lambda_a e_a,
  \qquad |\lambda_a|>1 .
\]
For the linear RG map
\[
  \mathcal R_{\rm lin}(u,s)=(Au,Bs),
\]
the unstable component of an initial condition
\[
  u_0=\sum_{a=1}^r u_a(0)e_a
\]
stays in the coordinate box
\begin{equation}
  |u_a(n)|\leq \rho
  \qquad (a=1,\ldots,r,\ 0\leq n\leq N)
  \label{eq:finite-scale-unstable-box}
\end{equation}
if and only if
\begin{equation}
  |u_a(0)|\leq \rho\,|\lambda_a|^{-N}
  \qquad (a=1,\ldots,r).
  \label{eq:finite-scale-tuning-precision}
\end{equation}
Writing \(\lambda_a=L^{y_a}\), this finite-depth accuracy is
\(\rho L^{-Ny_a}\) in the \(a\)-th relevant coordinate.

The linear recurrence gives
\[
  u_n=A^n u_0
  =
  \sum_{a=1}^r \lambda_a^n u_a(0)e_a .
\]
Therefore
\[
  |u_a(n)|=|\lambda_a|^n |u_a(0)|.
\]
Since \(|\lambda_a|>1\), the maximum over \(0\leq n\leq N\) occurs at
\(n=N\).  The bound \eqref{eq:finite-scale-unstable-box} is therefore
equivalent to
\[
  |\lambda_a|^N |u_a(0)|\leq\rho
\]
for every \(a\), which is
\eqref{eq:finite-scale-tuning-precision}.  The notation
\(\lambda_a=L^{y_a}\) gives \(|\lambda_a|^{-N}=L^{-Ny_a}\) when
\(\lambda_a>0\); for complex or negative eigenvalues the same estimate uses
\(|\lambda_a|=L^{y_a}\).

For a nonlinear RG map,
the finite-scale tuning calculation above is the linear model for the
precision required before \(N\) RG steps.  The stable graph
replaces the coordinate plane \(u=0\) by the intrinsic equation
\(\tau(V)=0\).  A theorem assigning the same exponents beyond the linear
approximation must provide normal-form or graph-transform estimates
controlling the nonlinear remainder.

\section{Polymer Representation}

On a lattice, a polymer is a finite connected union of blocks in a fixed block
decomposition.  Let \(\mathbb B_k\) be the set of scale-\(k\) blocks, let
\(\mathcal P_k\) be the family of connected finite unions of elements of
\(\mathbb B_k\), and let \(X\sim Y\) mean that the polymers \(X,Y\in
\mathcal P_k\) are disjoint and separated by at least one block.  A polymer
activity is a map
\[
  K_k:\mathcal P_k\times\Phi_k\to\mathbb C,\qquad
  (X,\phi)\mapsto K_k(X,\phi),
\]
where \(\Phi_k\) is the finite-dimensional scale-\(k\) field space of the
regulated model.  The polymer form of a finite-volume partition function is
\begin{equation}
  Z_k
  =
  \int d\mu_{C_k}(\phi)\,\exp\{-V_k(\phi)\}
  \sum_{\mathcal X\ {\rm compatible}}
  \prod_{X\in\mathcal X}K_k(X,\phi).
  \label{eq:polymer-partition-function-rg}
\end{equation}
Here \(V_k\) is the local part, \(C_k\) is the fluctuation covariance assigned
to the current scale, and the sum is over finite compatible families
\(\mathcal X\subset\mathcal P_k\).  Compatibility is the locality input: it
is what permits a disconnected family of small activities to exponentiate
through the cluster expansion rather than behaving as one large nonlocal
object.

\begin{definition}[Finite-range fluctuation covariance decomposition]
\label{def:finite-range-fluctuation-covariance-datum}
Let \(\Lambda_k\) be the finite scale-\(k\) regulator lattice with graph
distance \(d_k\).  A fluctuation covariance \(\Gamma_k\) has range
\(R_k\) if
\begin{equation}
  \Gamma_k(x,y)=0
  \qquad{\rm whenever}\qquad d_k(x,y)>R_k .
  \label{eq:finite-range-covariance-support}
\end{equation}
A finite-range decomposition for a scalar block-spin RG is a decomposition
\[
  C_{\leq k}=\Gamma_0+\Gamma_1+\cdots+\Gamma_k+C_{\rm rem,k}
\]
on each finite regulator, with every \(\Gamma_j\) positive semidefinite,
\(\Gamma_j\) satisfying a range bound of order \(L^j\) in microscopic
distance, and derivative or difference bounds in the field norm used by the
polymer activity.  Positivity is needed to realize \(\Gamma_j\) as a Gaussian
fluctuation covariance; the range bound is the input that turns separated
polymer activities into independent fluctuation integrations; the derivative
bounds enter the localization and Taylor-remainder estimates.
\end{definition}

\begin{lemma}[Finite-range Gaussian factorization]
\label{lem:finite-range-gaussian-factorization}
Let \(\Phi_\Lambda\) be a finite real field space and let
\(\mu_{\Gamma}\) be a centered Gaussian measure with positive semidefinite
covariance \(\Gamma\).  If \(A,B\subset\Lambda\) satisfy
\[
  \Gamma(x,y)=0\qquad (x\in A,\ y\in B),
\]
then the Gaussian fields \(\zeta_A=(\zeta_x)_{x\in A}\) and
\(\zeta_B=(\zeta_y)_{y\in B}\) are independent.  Consequently, for bounded
measurable \(F_A\) and \(F_B\) depending only on the fields in \(A\) and
\(B\),
\begin{equation}
  \int F_A(\zeta_A)F_B(\zeta_B)\,\dd\mu_\Gamma(\zeta)
  =
  \left(\int F_A(\zeta_A)\,\dd\mu_\Gamma(\zeta)\right)
  \left(\int F_B(\zeta_B)\,\dd\mu_\Gamma(\zeta)\right).
  \label{eq:finite-range-gaussian-factorization}
\end{equation}
\end{lemma}

\begin{proof}
The joint characteristic function of \((\zeta_A,\zeta_B)\) is
\[
  \exp\!\left\{
  -\frac12 t_A\Gamma_{AA}t_A
  -t_A\Gamma_{AB}t_B
  -\frac12 t_B\Gamma_{BB}t_B
  \right\}.
\]
The hypothesis gives \(\Gamma_{AB}=0\), so the characteristic function is the
product of the characteristic functions of \(\zeta_A\) and \(\zeta_B\).  For
finite-dimensional probability measures, characteristic functions determine
the law, and this factorization is exactly independence of the two random
vectors.  The displayed integral identity is then the defining factorization
property of independent random variables.  If \(\Gamma\) is only positive
semidefinite, the same argument applies to the induced Gaussian law on the
quotient by the null directions, or equivalently after adding
\(\epsilon 1\) and letting \(\epsilon\downarrow0\) in finite dimension.
\end{proof}

Applied with \(A\) and \(B\) equal to \(R_k\)-neighborhoods of two separated
polymers, Lemma~\ref{lem:finite-range-gaussian-factorization} is the exact
finite-regulator reason why fluctuation integration preserves polymer
locality.  If a covariance decomposition has only exponential decay rather
than finite range, this factorization is replaced by a cluster estimate with
decay constants in the polymer norm; the conclusion is weaker and must be
proved separately.  Thus a short-range scalar Wilsonian RG theorem cannot
begin from the phrase ``integrate a momentum shell'' alone.  It must specify
the covariance shell, its locality or decay estimate, and the norm in which
the residual inter-polymer correlations are summable.

\begin{lemma}[Summable covariance-tail bridge estimate]
\label{lem:summable-covariance-tail-bridge-estimate}
Let \(\Lambda\) be a finite lattice with graph distance \(d\).  Assume that
there are constants \(D_{\rm vol}\geq1\) and \(C_{\rm sh}<\infty\) such that
for every \(x\in\Lambda\) and \(n\geq0\)
\begin{equation}
  \#\{y\in\Lambda:\ d(x,y)=n\}
  \leq
  C_{\rm sh}(1+n)^{D_{\rm vol}-1}.
  \label{eq:rg-shell-count-bound}
\end{equation}
Let \(\Gamma\) be a positive semidefinite covariance satisfying
\begin{equation}
  |\Gamma(x,y)|
  \leq
  A_\Gamma(1+d(x,y))^{-s},
  \qquad s>D_{\rm vol}.
  \label{eq:rg-covariance-polynomial-tail-bound}
\end{equation}
If \(A,B\subset\Lambda\) are finite sets with \(d(A,B)\geq R\), then the
cross block \(\Gamma_{AB}:\ell^2(B)\to\ell^2(A)\) obeys
\begin{equation}
  \|\Gamma_{AB}\|_{\ell^2(B)\to\ell^2(A)}
  \leq
  A_\Gamma C_{\rm sh}
  \left(1+\frac{1}{s-D_{\rm vol}}\right)
  (1+R)^{D_{\rm vol}-s}.
  \label{eq:rg-covariance-tail-schur-bound}
\end{equation}
Consequently, if \(F_A\) and \(F_B\) are \(C^1\) bounded functions depending
only on the fields in \(A\) and \(B\), with
\[
  \|\nabla_A F_A\|_{2,\infty}
  =
  \sup_\phi
  \left(\sum_{x\in A}|\partial_xF_A(\phi)|^2\right)^{1/2}
\]
and similarly for \(F_B\), then
\begin{equation}
\begin{aligned}
  &\left|
    \mathbb E_\Gamma[F_A(\zeta_A)F_B(\zeta_B)]
    -
    \mathbb E_\Gamma[F_A(\zeta_A)]\,
    \mathbb E_\Gamma[F_B(\zeta_B)]
  \right|                                                    \\
  &\qquad\leq
  A_\Gamma C_{\rm sh}
  \left(1+\frac{1}{s-D_{\rm vol}}\right)
  (1+R)^{D_{\rm vol}-s}
  \|\nabla_AF_A\|_{2,\infty}\|\nabla_BF_B\|_{2,\infty}.
\end{aligned}
  \label{eq:rg-covariance-tail-connected-bound}
\end{equation}
\end{lemma}

\begin{proof}
For \(x\in A\), the row sum over \(B\) is bounded by the full tail outside
distance \(R\):
\[
  \sum_{y\in B}|\Gamma(x,y)|
  \leq
  A_\Gamma C_{\rm sh}
  \sum_{n\geq R}(1+n)^{D_{\rm vol}-1-s}.
\]
The same bound holds for column sums by symmetry of \(\Gamma\).  Since
\(s>D_{\rm vol}\),
\[
  \sum_{n\geq R}(1+n)^{D_{\rm vol}-1-s}
  \leq
  (1+R)^{D_{\rm vol}-1-s}
  +
  \int_R^\infty (1+t)^{D_{\rm vol}-1-s}\,\dd t
  \leq
  \left(1+\frac1{s-D_{\rm vol}}\right)
  (1+R)^{D_{\rm vol}-s}.
\]
Schur's test gives
\eqref{eq:rg-covariance-tail-schur-bound}.

It remains to connect the operator bound to the Gaussian expectation.  For
\(0\leq t\leq1\), let \(\mu_t\) be the centered Gaussian law on
\(\mathbb R^A\oplus\mathbb R^B\) with covariance
\[
  \begin{pmatrix}
    \Gamma_{AA} & t\Gamma_{AB}\\
    t\Gamma_{BA} & \Gamma_{BB}
  \end{pmatrix}.
\]
The covariance is positive because it is obtained from the original Gaussian
law by multiplying the \(B\)-component by an independent sign and averaging:
\(t=2p-1\) for \(p=(1+t)/2\).  Differentiating the finite-dimensional
Gaussian expectation with respect to \(t\) gives
\[
  \frac{\dd}{\dd t}\mathbb E_{\mu_t}[F_A F_B]
  =
  \sum_{x\in A,y\in B}\Gamma(x,y)
  \mathbb E_{\mu_t}[(\partial_xF_A)(\partial_yF_B)] .
\]
At \(t=0\) the \(A\) and \(B\) variables are independent; at \(t=1\) the law
is the original marginal of \(\mu_\Gamma\) on \(A\cup B\).  Integrating in
\(t\), applying Cauchy--Schwarz in the finite index sets, and using
\eqref{eq:rg-covariance-tail-schur-bound} gives
\eqref{eq:rg-covariance-tail-connected-bound}.
\end{proof}

Lemma~\ref{lem:summable-covariance-tail-bridge-estimate} is the replacement
for exact factorization when the covariance shell is not finite range.  The
decay exponent, shell-count constant, derivative norm, and polymer separation
enter the RG theorem as actual constants.  A smooth momentum-shell covariance
can supply such constants after integration by parts in momentum space; a
sharp cutoff or a long-range covariance may fail to give the summability
needed by the chosen polymer norm.  This is why a rigorous Wilsonian proof
states the covariance decomposition and cluster norm before using locality
language.

\begin{lemma}[Finite localization remainder and scaling gain]
\label{lem:finite-localization-remainder-scaling-gain}
Let \(E\) be a finite-dimensional normed real vector space, let
\(F:E\to\mathbb R\) be \(C^{p+1}\) on the closed ball
\(\{\xi:\|\xi\|\leq r\}\), and let
\[
  (\mathcal L_{\leq p}F)(\xi)
  =
  \sum_{n=0}^{p}\frac{1}{n!}D^nF(0)[\xi,\ldots,\xi]
\]
be the Taylor-localization map at the origin.  If
\[
  M_{p+1}(r)
  =
  \sup_{\|\eta\|\leq r}\|D^{p+1}F(\eta)\|_{\rm op},
\]
then, for every \(\|\xi\|\leq r\),
\begin{equation}
  |F(\xi)-(\mathcal L_{\leq p}F)(\xi)|
  \leq
  \frac{M_{p+1}(r)}{(p+1)!}\,\|\xi\|^{p+1}.
  \label{eq:finite-taylor-localization-remainder-bound}
\end{equation}
In a scalar block-spin chart with scale factor \(L\), dimension \(D\), field
scaling dimension \(\Delta_\phi\), and one lattice difference carrying one
inverse power of \(L\), a local density monomial with \(\nu\) fields and
\(m\) lattice differences has canonical scaling exponent
\begin{equation}
  y_{\nu,m}=D-\nu\Delta_\phi-m .
  \label{eq:canonical-localization-scaling-exponent}
\end{equation}
Thus a Taylor-localization choice gives a linear irrelevant gain
\(L^{-\alpha}\) only after the first omitted local monomials satisfy
\begin{equation}
  \alpha
  =
  \inf_{\text{omitted }(\nu,m)}
  \bigl(\nu\Delta_\phi+m-D\bigr)
  >0 .
  \label{eq:first-omitted-localization-gap}
\end{equation}
\end{lemma}

\begin{proof}
The integral Taylor formula in finite dimension gives
\[
  F(\xi)-(\mathcal L_{\leq p}F)(\xi)
  =
  \frac{1}{p!}
  \int_0^1
    (1-s)^p D^{p+1}F(s\xi)[\xi,\ldots,\xi]\,\dd s .
\]
Taking the operator norm and using
\(\int_0^1(1-s)^p\dd s=(p+1)^{-1}\) gives
\eqref{eq:finite-taylor-localization-remainder-bound}.  For the scaling
formula, write a fine field inside one coarse block as
\(\phi=L^{-\Delta_\phi}\phi'\).  The sum or integral over a coarse block
contains \(L^D\) fine cells, while each discrete difference contributes an
additional factor \(L^{-1}\).  Hence the coefficient of a monomial with
\(\nu\) fields and \(m\) differences is multiplied, at the canonical
linearized level, by
\[
  L^D L^{-\nu\Delta_\phi}L^{-m}=L^{D-\nu\Delta_\phi-m}.
\]
The first omitted local monomial is contractive by a factor at most
\(L^{-\alpha}\) exactly when the exponents of all omitted monomials are
\(\leq-\alpha\), which is
\eqref{eq:first-omitted-localization-gap}.
\end{proof}

This lemma is finite-regulator bookkeeping, but it removes an important
ambiguity.  The localization map \(\mathcal L_k\) in a polymer RG proof is
a Taylor jet with a declared center, derivative norm, and retained local
coordinate set.  It must be paired with the derivative norm in
\eqref{eq:finite-taylor-localization-remainder-bound}, the field rescaling
used in the chart, and a proof that the omitted Taylor coordinates are
contractive in that chart.  In canonical scalar bookkeeping at \(D=4\),
\(\Delta_\phi=1\) makes \(\phi^4\) marginal and \(\phi^6\) irrelevant by two
powers.  In canonical bookkeeping at \(D=3\),
\(\Delta_\phi=\frac12\) makes the \(\phi^6\) density marginal by
\eqref{eq:canonical-localization-scaling-exponent}; it cannot be discarded
into the irrelevant remainder on engineering grounds alone.  An interacting
fixed-point proof may replace the canonical monomial coordinate list by scaling-field
coordinates, but that replacement is itself part of the model-specific RG
theorem.

\begin{lemma}[Finite Gaussian stability of a quadratic large-field regulator]
\label{lem:finite-gaussian-large-field-regulator-stability}
Let \(E\) be a finite-dimensional real Hilbert space, let
\(\Gamma:E\to E\) be positive semidefinite and self-adjoint, and let
\(\mu_\Gamma\) be the centered Gaussian law with covariance \(\Gamma\).
For \(\kappa>0\) assume
\[
  2\kappa\|\Gamma\|_{\rm op}<1 .
\]
Then, for every \(\phi\in E\),
\begin{equation}
  \int_E
    \exp\{\kappa\|\phi+\zeta\|^2\}\,\dd\mu_\Gamma(\zeta)
  =
  \det(1-2\kappa\Gamma)^{-1/2}
  \exp\{\kappa\langle\phi,(1-2\kappa\Gamma)^{-1}\phi\rangle\}.
  \label{eq:gaussian-large-field-regulator-exact}
\end{equation}
Consequently
\begin{equation}
  \int_E
    \exp\{\kappa\|\phi+\zeta\|^2\}\,\dd\mu_\Gamma(\zeta)
  \leq
  C_\Gamma(\kappa)\,
  \exp\!\left\{
    \frac{\kappa}{1-2\kappa\|\Gamma\|_{\rm op}}\|\phi\|^2
  \right\},
  \label{eq:gaussian-large-field-regulator-bound}
\end{equation}
where \(C_\Gamma(\kappa)=\det(1-2\kappa\Gamma)^{-1/2}\) on
\((\ker\Gamma)^\perp\).
\end{lemma}

\begin{proof}
It is enough to work on \((\ker\Gamma)^\perp\).  Choose an orthonormal basis
diagonalizing \(\Gamma\), with eigenvalues \(\lambda_i>0\).  The integral is
a product of one-dimensional integrals
\[
  \frac{1}{\sqrt{2\pi\lambda_i}}
  \int_{\mathbb R}
    \exp\!\left\{
      \kappa(\phi_i+\zeta_i)^2-\frac{\zeta_i^2}{2\lambda_i}
    \right\}\dd\zeta_i .
\]
Since \(2\kappa\lambda_i<1\), completing the square gives
\[
  \kappa(\phi_i+\zeta_i)^2-\frac{\zeta_i^2}{2\lambda_i}
  =
  -\left(\frac{1}{2\lambda_i}-\kappa\right)
    \left(\zeta_i-\frac{\kappa\phi_i}
      {\frac{1}{2\lambda_i}-\kappa}\right)^2
  +
  \frac{\kappa}{1-2\kappa\lambda_i}\phi_i^2 .
\]
The Gaussian integral in the shifted variable contributes
\((1-2\kappa\lambda_i)^{-1/2}\).  Multiplying over \(i\) gives
\eqref{eq:gaussian-large-field-regulator-exact}.  Since
\((1-2\kappa\lambda_i)^{-1}\leq
(1-2\kappa\|\Gamma\|_{\rm op})^{-1}\), the quadratic form in the exponent is
bounded by the exponent in
\eqref{eq:gaussian-large-field-regulator-bound}.
\end{proof}

Thus a large-field regulator such as
\(\exp\{\kappa\|\phi\|^2_X\}\) is stable under fluctuation integration only
with a declared covariance spectral bound and with the next-scale regulator
large enough to absorb the enlarged coefficient
\(\kappa(1-2\kappa\|\Gamma\|_{\rm op})^{-1}\).  In a polymer RG theorem this
finite estimate is combined with finite range or cluster decay and with
Taylor localization.  The constants \(C_\Gamma(\kappa)\), the covariance
spectral bound, and the regulator strengthening are part of the proof data
entering \(A_{\rm irr}\), \(B_{\rm pol}\), and the extraction defect.

\begin{definition}[One-step polymer RG contraction estimates]
\label{def:one-step-polymer-rg-contraction-datum}
A one-step polymer RG contraction structure at scale \(k\) consists of the
following finite-regulator data.
\begin{enumerate}
  \item A scale factor \(L>1\), a reblocking map
  \(X\mapsto \overline X\) from scale-\(k\) polymers to the smallest
  scale-\(k+1\) polymer containing \(X\), and a finite overlap constant
  \begin{equation}
    \mathfrak C_a
    =
    \sup_{\Delta\in\mathbb B_{k+1}}
    \sum_{\substack{X\in\mathcal P_k\\ \overline X\supset\Delta}}
    \exp\{-a|X|_k\}
    <\infty .
    \label{eq:polymer-overlap-constant}
  \end{equation}
  \item Large-field regulators \(G_k(X,\phi)\geq1\), analytic radii \(h_k\),
  and a weighted polymer norm
  \begin{equation}
    \|K_k\|_{k,a,h}
    =
    \sup_{\Delta\in\mathbb B_k}
    \sum_{\substack{X\in\mathcal P_k\\ X\supset\Delta}}
    \exp\{a|X|_k\}
    \sup_{\phi\in\Phi_k}
    \frac{|K_k(X,\phi)|_{h_k}}{G_k(X,\phi)} .
    \label{eq:rg-polymer-norm}
  \end{equation}
  The seminorm \(|\cdot|_{h_k}\) is the finite Taylor or analytic norm in the
  field variables used by the model.  Its precise definition is part of the
  contraction structure, because the contraction constant depends on the allowed derivatives
  and on the large-field weights.
  \item A localization map
  \(\mathcal L_k\) from polymer functionals supported in one reblocked
  polymer to the chosen finite-dimensional local-coordinate space.  The
  complementary projection \(1-\mathcal L_k\) is the irrelevant remainder.
  After fluctuation integration and rescaling, denoted here by
  \(\mathcal A_k\), the contraction structure includes constants
  \(A_{\rm irr}<\infty\) and \(\alpha>0\) such that
  \begin{equation}
    \|(1-\mathcal L_k)\mathcal A_kK_k\|_{k+1,a,h}
    \leq
    A_{\rm irr}L^{-\alpha}\|K_k\|_{k,a,h}.
    \label{eq:rg-polymer-linear-irrelevant-bound}
  \end{equation}
  This is the analytic place where Taylor remainder estimates, covariance
  finite-range bounds, field rescaling, and large-field stability enter.
  \item A quadratic product estimate for incompatible small polymers.  For
  two activities define the scale-\(k+1\) circle product by
  \begin{equation}
    (K_1\circ_k K_2)(Y,\phi)
    =
    \sum_{\substack{X_1,X_2\in\mathcal P_k\\
      \overline{X_1\cup X_2}=Y\\ X_1\not\sim X_2}}
    K_1(X_1,\phi)\,K_2(X_2,\phi),
    \label{eq:rg-polymer-circle-product}
  \end{equation}
  with the product understood after pulling both activities to the same
  reblocked field variables.  The contraction structure requires a constant
  \(B_{\rm pol}<\infty\) such that
  \begin{equation}
    \|K_1\circ_k K_2\|_{k+1,a,h}
    \leq
    B_{\rm pol}\|K_1\|_{k,a,h}\|K_2\|_{k,a,h}.
    \label{eq:rg-polymer-circle-product-bound}
  \end{equation}
  The constant \(B_{\rm pol}\) is a consequence of the finite overlap
  constant \(\mathfrak C_a\), the compatibility convention, and the
  multiplicative behavior of \(G_k\).  It is not a universal number.

  The finite-overlap content can be made explicit at the level of scalar
  majorants.  Put
  \[
    M_i(X)
    =
    \sup_{\phi\in\Phi_k}
    \frac{|K_i(X,\phi)|_{h_k}}{G_k(X,\phi)} .
  \]
  Suppose that the analytic norm and large-field regulator satisfy the
  submultiplicative estimate
  \begin{equation}
    \frac{|K_1(X_1,\phi)K_2(X_2,\phi)|_{h_{k+1}}}
         {G_{k+1}(\overline{X_1\cup X_2},\phi)}
    \leq
    C_G\,M_1(X_1)M_2(X_2)
    \label{eq:polymer-regulator-submultiplicative-bound}
  \end{equation}
  for every incompatible pair entering
  \eqref{eq:rg-polymer-circle-product}.  Define the pair-overlap majorant
  \begin{equation}
    \mathfrak C_a^\circ
    =
    \sup_{\Delta\in\mathbb B_{k+1}}
    \sum_{\substack{X_1,X_2\in\mathcal P_k\\
      X_1\not\sim X_2,
      \overline{X_1\cup X_2}\supset\Delta}}
    \exp\!\left\{
      a|\overline{X_1\cup X_2}|_{k+1}
      -a|X_1|_k-a|X_2|_k
    \right\}.
    \label{eq:polymer-pair-overlap-majorant}
  \end{equation}
  If \(\mathfrak C_a^\circ<\infty\), then
  \(B_{\rm pol}=C_G\mathfrak C_a^\circ\) is admissible in
  \eqref{eq:rg-polymer-circle-product-bound}.  Indeed, from
  \eqref{eq:rg-polymer-norm},
  \[
    M_i(X_i)
    \leq
    e^{-a|X_i|_k}\|K_i\|_{k,a,h}
  \]
  because the norm sum contains the term \(X_i\) for each block contained in
  \(X_i\).  Therefore, for each coarse block \(\Delta\),
  \[
  \begin{aligned}
    &\sum_{Y\supset\Delta}e^{a|Y|_{k+1}}
      \sup_\phi
      \frac{|(K_1\circ_kK_2)(Y,\phi)|_{h_{k+1}}}{G_{k+1}(Y,\phi)}
      \\
    &\qquad\leq
      C_G\|K_1\|_{k,a,h}\|K_2\|_{k,a,h}
      \sum_{\substack{X_1,X_2\\
      X_1\not\sim X_2,
      \overline{X_1\cup X_2}\supset\Delta}}
      e^{a|\overline{X_1\cup X_2}|_{k+1}
        -a|X_1|_k-a|X_2|_k}
      \\
    &\qquad\leq
      C_G\mathfrak C_a^\circ
      \|K_1\|_{k,a,h}\|K_2\|_{k,a,h}.
  \end{aligned}
  \]
  The model-specific proof must bound \(\mathfrak C_a^\circ\) uniformly in
  the volume and scale.  On a finite-range block lattice this is a geometric
  counting estimate; with long-range covariances or nonlocal blockings it may
  fail unless the polymer norm is strengthened.
  \item A local-coordinate extraction error \(\varepsilon_k\geq0\), measured
  in the same polymer norm, accounting for the difference between the exact
  finite-regulator RG step and the chosen local chart.  In a fully exact
  chart \(\varepsilon_k=0\); in a truncated chart it is a theorem-level
  remainder and must be estimated.
\end{enumerate}
\end{definition}

The preceding datum turns contraction of the irrelevant activity into a
finite inequality.  If
\[
  x_k=\|K_k\|_{k,a,h},\qquad
  q=A_{\rm irr}L^{-\alpha},
\]
then the linear estimate
\eqref{eq:rg-polymer-linear-irrelevant-bound}, the quadratic product estimate
\eqref{eq:rg-polymer-circle-product-bound}, and the extraction error give the
one-step bound
\begin{equation}
  x_{k+1}
  \leq
  qx_k+B_{\rm pol}x_k^2+\varepsilon_k .
  \label{eq:rg-polymer-recursion-budget}
\end{equation}
Consequently a closed radius-\(r\) polymer ball is invariant with contraction
margin \(\theta<1\) whenever
\begin{equation}
  q+B_{\rm pol}r+\frac{\varepsilon_k}{r}\leq \theta<1
  \qquad{\rm uniformly\ in\ }k .
  \label{eq:rg-polymer-contraction-smallness}
\end{equation}
Indeed, \(x_k\leq r\) gives
\[
  x_{k+1}
  \leq
  \left(q+B_{\rm pol}r+\frac{\varepsilon_k}{r}\right)r
  \leq \theta r .
\]
This is elementary arithmetic, but it is the arithmetic that fixes the
meaning of the polymer estimate: every proof must exhibit the norm, the
large-field regulator, the localization operator, the finite overlap bound,
and the uniform smallness inequality.  Without these data the displayed
contraction is a slogan, not a Wilsonian RG theorem.

The same data must be controlled as a sequence of scales, not only at one
scale.  Suppose that, for \(0\leq j<n\), the trajectory stays in the
radius-\(r\) polymer ball and
\[
  q+B_{\rm pol}r\leq\theta<1 .
\]
Then \eqref{eq:rg-polymer-recursion-budget} implies the forced linear
recursion
\[
  x_{j+1}\leq \theta x_j+\varepsilon_j
  \qquad (0\leq j<n),
\]
and iteration gives
\begin{equation}
  x_n
  \leq
  \theta^n x_0+
  \sum_{j=0}^{n-1}\theta^{\,n-1-j}\varepsilon_j .
  \label{eq:rg-polymer-multiscale-forcing-bound}
\end{equation}
Thus a scale-dependent extraction error is harmless only after its weighted
sum is controlled.  If
\(\varepsilon_j\leq E\sigma^j\) with \(0<\sigma<1\), then
\begin{equation}
  \sum_{j=0}^{n-1}\theta^{\,n-1-j}\varepsilon_j
  \leq
  E\frac{\theta^n-\sigma^n}{\theta-\sigma}
  \quad(\theta\neq\sigma),
  \qquad
  \leq En\theta^{n-1}
  \quad(\theta=\sigma).
  \label{eq:rg-polymer-geometric-forcing-bound}
\end{equation}
For a nondecaying uniform defect \(\varepsilon_j\leq E\), the corresponding
bound is
\[
  x_n\leq\theta^nx_0+\frac{E(1-\theta^n)}{1-\theta}.
\]
Consequently the ball of radius \(r\) is invariant for all scales under the
stronger smallness condition
\begin{equation}
  x_0+\frac{E}{1-\theta}\leq r
  \label{eq:rg-polymer-uniform-forcing-invariant-ball}
\end{equation}
in the uniform-forcing case, or under the analogous supremum of the weighted
geometric sum in the decaying case.  These estimates are still only
bookkeeping; the model-specific constructive theorem must prove the bounds
on \(q\), \(B_{\rm pol}\), \(r\), and \(\varepsilon_j\) from the covariance,
localization, and large-field regulators.

\begin{lemma}[Finite source-window extraction and propagated cumulant error]
\label{lem:finite-source-window-extraction-error}
Let \(B_j\) be Banach spaces and let \(R_j(J)\in B_j\) be holomorphic on a
neighborhood of the closed polydisc
\[
  P_{\rho_j}
  =
  \{J\in\mathbb C^m:\ |J_a|\leq \rho_{j,a},\ 1\leq a\leq m\}.
\]
Set
\[
  \eta_j=\sup_{J\in P_{\rho_j}}\|R_j(J)\|_{B_j}.
\]
For an integer \(s_j\geq0\), let \(\mathcal T_{\leq s_j}^{\rm src}R_j\) be
the Taylor polynomial in \(J\) containing all monomials of total source
degree at most \(s_j\).  If \(\beta=(\beta_1,\ldots,\beta_m)\) is a
multi-index and
\[
  \rho_j^\beta=\prod_{a=1}^{m}\rho_{j,a}^{\beta_a},
  \qquad
  \beta!=\prod_{a=1}^{m}\beta_a!,
\]
then
\begin{equation}
  \partial_J^\beta
  \bigl(R_j-\mathcal T_{\leq s_j}^{\rm src}R_j\bigr)(0)=0
  \qquad (|\beta|\leq s_j),
  \label{eq:source-window-retained-derivatives-vanish}
\end{equation}
and, for arbitrary \(\beta\),
\begin{equation}
  \left\|
  \partial_J^\beta
  \bigl(R_j-\mathcal T_{\leq s_j}^{\rm src}R_j\bigr)(0)
  \right\|_{B_j}
  \leq
  \frac{\beta!}{\rho_j^\beta}\,\eta_j .
  \label{eq:source-window-cauchy-bound}
\end{equation}
Suppose moreover that the error in a reconstructed connected source window
at scale \(N\) is obtained by propagating the source-extraction tails through
bounded linear maps
\[
  U_{j\to N,\beta}:B_j\longrightarrow \mathcal O_\beta,
  \qquad
  p_\beta(U_{j\to N,\beta}v)
  \leq
  P_{j,N,\beta}\|v\|_{B_j},
\]
where \(p_\beta\) is the seminorm of the declared \(\beta\)-source
observable window.  Then the accumulated \(\beta\)-window error satisfies
\begin{equation}
  p_\beta(\Delta O_{N,\beta})
  \leq
  \sum_{\substack{0\leq j<N\\ s_j<|\beta|}}
  P_{j,N,\beta}\,
  \frac{\beta!}{\rho_j^\beta}\,
  \eta_j .
  \label{eq:source-window-propagated-error-bound}
\end{equation}
\end{lemma}

\begin{proof}
The Taylor polynomial contains exactly the source monomials of total degree
at most \(s_j\), hence the derivative in
\eqref{eq:source-window-retained-derivatives-vanish} vanishes for
\(|\beta|\leq s_j\).  For \(|\beta|>s_j\), the \(\beta\)-derivative of the
tail at the origin is the \(\beta\)-derivative of \(R_j\).  Use the
Banach-valued Cauchy integral formula on the distinguished boundary
\(|Z_a|=\rho_{j,a}\):
\[
  \partial_J^\beta R_j(0)
  =
  \frac{\beta!}{(2\pi i)^m}
  \int_{|Z_1|=\rho_{j,1}}\cdots
  \int_{|Z_m|=\rho_{j,m}}
  \frac{R_j(Z)}{Z^{\beta+1}}
  \,\dd Z_1\cdots\dd Z_m .
\]
Taking the \(B_j\)-norm and using the boundary length factors gives
\eqref{eq:source-window-cauchy-bound}.  The propagated estimate follows by
applying \(U_{j\to N,\beta}\) to each scale-\(j\) source-tail derivative,
using its operator bound, and summing over scales.  Terms with
\(s_j\geq|\beta|\) vanish by
\eqref{eq:source-window-retained-derivatives-vanish}.
\end{proof}

Thus a source-extended polymer RG chart must declare which source monomials
are retained and which source derivatives of the connected generating
functional are being reconstructed.  To control all \(n\)-point Schwinger
distribution windows in a fixed test-function family, the source-coordinate system must
retain source degree at least \(n\), or it must prove the Cauchy-tail sum
\eqref{eq:source-window-propagated-error-bound}.  A scalar fixed-point
trajectory without this source-extension estimate is a theorem about the
partition function or un-sourced activity only; it is not yet a theorem
about the reconstructed local QFT correlators.

For ordinary short-range critical scalar models, the unresolved work is to
construct this datum with constants uniform along the tuned critical
trajectory and with source extensions strong enough for reconstructed
correlation functions.  For gauge theories, the same structure must be built
on gauge-invariant lattice activities or in a BV-compatible finite-regulator
chart; a gauge-fixed formal activity without the quantum master equation or a
gauge-invariant blocking map is not a Wilsonian gauge-theory construction in
the sense used here.

\section{Fixed Points And Continuum Limits}

A continuum limit is obtained by tuning the unstable coordinates so that the
RG trajectory remains near a fixed point for arbitrarily many steps and then
converges, after reconstruction, to regulator-independent correlation
functions.  The existence of a non-Gaussian fixed point as a Banach-space
fixed point is a theorem only after the RG map, norm, domain, and estimates
are supplied.  Perturbative beta functions and numerical fixed-point searches
are evidence for a coordinate chart; they do not replace the fixed-point
construction.

\begin{definition}[Theorem-level nonperturbative RG fixed point]
\label{def:theorem-level-nonperturbative-rg-fixed-point}
A nonperturbative Wilsonian fixed-point theorem consists of the following
declared data.
\begin{enumerate}
  \item A regulator class and a concrete RG map
  \(\mathcal R:\mathcal U\to\mathcal B\) on an open subset
  \(\mathcal U\) of a Banach space \((\mathcal B,\|\cdot\|)\) of
  interactions, measures, tensors, or polymer activities.
  \item A proof that the map is defined on the claimed domain and is
  differentiable to the order used in the argument.
  \item A point \(V_\ast\in\mathcal U\) satisfying
  \(\mathcal R(V_\ast)=V_\ast\), or a periodic orbit when the intended RG
  structure is not a fixed point.
  \item A spectral or graph-transform analysis of
  \(D\mathcal R|_{V_\ast}\), including the relevant, marginal, irrelevant, and
  redundant subspaces in the chosen norm.
  \item A reconstruction map from the tuned sequence of finite-cutoff
  theories to correlation functions, local algebras, or Euclidean
  Schwinger functions, with convergence estimates sufficient for the claimed
  QFT framework.
\end{enumerate}
The word ``fixed point'' in a theorem of this kind refers to the specified
map \(\mathcal R\) and norm.  A zero of a truncated beta function, a solution
of a finite-dimensional functional-RG truncation, or a numerical tensor-RG
fixed point is evidence for such a theorem only after an embedding into this
list has been supplied.
\end{definition}

\begin{theorem}[Validated fixed point by a Newton--Kantorovich estimate]
\label{thm:newton-kantorovich-rg-fixed-point}
Let \(\mathcal B\) be a Banach space, let \(F:\mathcal U\to\mathcal B\) be a
\(C^1\) map on an open ball \(B_R(V_0)\subset\mathcal U\), and let
\[
  F(V)=\mathcal R(V)-V .
\]
Assume that \(A:\mathcal B\to\mathcal B\) is bounded and injective, and that
there are constants \(\eta,\rho,L\geq0\) with
\begin{align}
  \|A F(V_0)\|&\leq \eta, \label{eq:nk-residual-bound}\\
  \|1-A\,DF(V_0)\|_{\mathcal B\to\mathcal B}&\leq \rho,\\
  \|A(DF(V)-DF(W))\|_{\mathcal B\to\mathcal B}
     &\leq L\|V-W\|
       \qquad (V,W\in B_R(V_0)).
       \label{eq:nk-lipschitz-bound}
\end{align}
If there is \(r\leq R\) such that
\begin{equation}
  \eta+\rho r+\frac{L}{2}r^2\leq r,
  \qquad
  \rho+Lr<1,
  \label{eq:nk-ball-condition}
\end{equation}
then \(F(V)=0\) has a unique solution \(V_\ast\) in the closed ball
\(\overline B_r(V_0)\).  Equivalently, \(\mathcal R(V_\ast)=V_\ast\).
\end{theorem}

\begin{proof}
Define the Newton-like map
\[
  \mathcal N(V)=V-AF(V).
\]
For \(V\in\overline B_r(V_0)\), the fundamental theorem of calculus gives
\[
  \mathcal N(V)-V_0
  =
  -AF(V_0)
  +
  \int_0^1
  \bigl(1-A\,DF(V_0+s(V-V_0))\bigr)(V-V_0)\,\dd s .
\]
Using \eqref{eq:nk-residual-bound} and
\eqref{eq:nk-lipschitz-bound}, the norm of the right-hand side is bounded by
\[
  \eta+\rho r+\frac{L}{2}r^2,
\]
which is at most \(r\) by \eqref{eq:nk-ball-condition}.  Thus
\(\mathcal N\) maps \(\overline B_r(V_0)\) into itself.  For
\(V,W\in\overline B_r(V_0)\),
\[
  \mathcal N(V)-\mathcal N(W)
  =
  \int_0^1
  \bigl(1-A\,DF(W+s(V-W))\bigr)(V-W)\,\dd s ,
\]
and the same estimates give
\[
  \|\mathcal N(V)-\mathcal N(W)\|
  \leq (\rho+Lr)\|V-W\|.
\]
The coefficient is strictly less than one, so Banach's fixed-point theorem
gives a unique fixed point of \(\mathcal N\) in the ball.  If
\(\mathcal N(V_\ast)=V_\ast\), then \(AF(V_\ast)=0\).  Since \(A\) is
injective, \(F(V_\ast)=0\).  Conversely any zero of \(F\) is a fixed point of
\(\mathcal N\), so uniqueness follows.
\end{proof}

This theorem isolates the final functional-analytic step after RG estimates
have been supplied.  The difficult part in constructive RG is to choose
\(\mathcal B\), \(V_0\), and \(A\), and then prove
\eqref{eq:nk-residual-bound}--\eqref{eq:nk-lipschitz-bound} with constants
small enough that \eqref{eq:nk-ball-condition} holds.  In polymer RG this
means summing tree or cluster expansions with large-field regulators; in
tensor RG it means proving that the disentangled coarse-graining map is a
contraction on a neighborhood that is stable under the infinite-dimensional
map.

\section{Projection Truncations And Residual Control}

Finite-dimensional RG calculations enter the monograph only through a
declared projection.  Let \(P_N:\mathcal B\to\mathcal B_N\) be a bounded
projection onto a finite-dimensional subspace and put \(Q_N=1-P_N\).  For
\[
  F(V)=\mathcal R(V)-V ,
\]
the projected fixed-point equation is
\begin{equation}
  P_NF(V_N)=0,\qquad V_N\in\mathcal B_N .
  \label{eq:projected-rg-fixed-point-equation}
\end{equation}
The complement residual is
\begin{equation}
  \epsilon_N(V_N;A)
  =
  \|A Q_NF(V_N)\|,
  \label{eq:rg-complement-residual}
\end{equation}
where \(A\) is the approximate inverse used in the
Newton--Kantorovich estimate.

\begin{example}[Spurious projected zeros]
Projection residuals must be controlled in the omitted directions.  The
finite-dimensional model
\(\mathcal B=\mathbb R^2\), \(P(x,y)=(x,0)\), and
\[
  F(x,y)=(0,1).
\]
has \(P F(x,0)=0\) for every \(x\in\mathbb R\), while the second component of
\(F(x,y)\) is identically one.  Thus the entire projected line solves
\eqref{eq:projected-rg-fixed-point-equation}, and no point of the full space
solves \(F(V)=0\).  The residual
\(\epsilon_N(V_N;A)=\|AQ_NF(V_N)\|\) is the datum that detects this failure in
the Newton--Kantorovich lifting problem.
\end{example}

\paragraph{Lifting a projected RG zero.}
Let \(V_N\in\mathcal B_N\) solve
\eqref{eq:projected-rg-fixed-point-equation}.  Suppose the hypotheses of
Theorem~\ref{thm:newton-kantorovich-rg-fixed-point} hold with \(V_0=V_N\).
Then a true fixed point \(V_\ast\) of \(\mathcal R\) exists in
\(\overline B_r(V_N)\) whenever
\begin{equation}
  \epsilon_N(V_N;A)+\rho r+\frac{L}{2}r^2\le r,
  \qquad
  \rho+Lr<1 .
  \label{eq:projected-rg-lift-condition}
\end{equation}
Since \(P_NF(V_N)=0\), the full residual is
\[
  F(V_N)=Q_NF(V_N).
\]
Therefore
\[
  \|AF(V_N)\|=\|AQ_NF(V_N)\|=\epsilon_N(V_N;A).
\]
The conditions in
\eqref{eq:projected-rg-lift-condition} are exactly the
Newton--Kantorovich conditions \eqref{eq:nk-ball-condition} with
\(\eta=\epsilon_N(V_N;A)\).  Theorem~\ref{thm:newton-kantorovich-rg-fixed-point}
then gives a unique zero of \(F\) in \(\overline B_r(V_N)\), equivalently a
fixed point of \(\mathcal R\).

The companion calculation
\[
  \texttt{calculation-checks/rg\_projection\_checks.py}
\]
checks these finite-dimensional algebraic examples with exact rational
arithmetic.  In an actual Wilsonian RG proof, the hard work is to bound
\(\epsilon_N(V_N;A)\), \(\rho\), and \(L\) in the infinite-dimensional norm,
not to solve the projected equation.

\section{Constructive Fixed-Point Output Data}

The most useful rigorous fixed-point examples produce a fixed interaction
together with correlation and response functions whose remainders are
controlled in a
declared norm.  The following output record is the minimum structure recorded in this
monograph when using the recent long-range fermionic RG constructions as a
benchmark.

\begin{definition}[Constructive fermionic fixed-point output]
\label{def:constructive-fermionic-fixed-point-output}
A constructive fermionic fixed-point output consists of:
\begin{enumerate}
  \item a finite ultraviolet cutoff and a Grassmann Gaussian covariance
  \(C_\Lambda\), including its scaling law;
  \item a Banach space \(\mathcal B_{\rm ker}\) of antisymmetric interaction
  kernels, with a norm controlling decay, differentiability, and scale
  weights;
  \item an exact RG map \(\mathcal R\) on an open subset of
  \(\mathcal B_{\rm ker}\), defined by Grassmann Gaussian integration,
  localization of relevant coordinates, and rescaling;
  \item a fixed interaction \(V_\ast\in\mathcal B_{\rm ker}\), satisfying
  \(\mathcal R(V_\ast)=V_\ast\);
  \item source coordinates for the field and for the density, with scaling
  dimensions \([\psi]\) and \([J]\);
  \item response functions constructed from the fixed-point generating
  functional, together with scale-covariance estimates and explicit
  remainders.
\end{enumerate}
This is fixed-cutoff constructive RG output.  A continuum-QFT theorem
requires an additional removal or interpretation of the cutoff in the
observable topology being claimed.
\end{definition}

\begin{quotedtheorem}[Long-range fermionic \(\psi^4_d\) fixed point]
\label{qthm:long-range-fermionic-psi4-fixed-point}
For the long-range fermionic \(\psi^4_d\) model in dimensions
\(d=1,2,3\), with fractional kinetic term chosen so that the quartic
interaction is weakly relevant, there is a small-parameter regime in which
the constructive RG map has a nontrivial fixed interaction \(V_\ast\) in the
Banach space of kernels.  With a fixed ultraviolet cutoff, the field and
density response functions at \(V_\ast\) are constructed by convergent tree
expansions.  The field exponent is the naive exponent, while the density
exponent is anomalous and analytic in the small parameter.  The associated
scaling operators have two-point functions that are scale covariant up to a
stretched-exponential large-distance remainder controlled by the ultraviolet
cutoff.
\end{quotedtheorem}

\paragraph{Status of the quoted theorem.}
This is the theorem-level output of the GMR and GMRS constructive RG works
just quoted.  Its proof constructs \(V_\ast\) in an infinite-dimensional
kernel space and controls the irrelevant tail by a convergent multiscale tree
expansion.  The theorem is quoted here as a state-of-the-art benchmark for
what a nonperturbative Wilsonian fixed-point statement must contain.  The
monograph uses the theorem to calibrate the form of a complete RG claim, while
separate scalar and gauge theory claims in this book require their own
estimates.

\paragraph{Kernel-space mechanism in the fermionic benchmark.}
The fermionic example supplies a constructive fixed point by building an
invariant infinite-dimensional kernel space and proving that the RG map closes
on it.  Write a source-extended monomial by a label
\[
  \alpha=(n,m,\ell,p),
  \qquad
  p=(p_1,\ldots,p_\ell),\quad p_i\in\{0,1\},
\]
where \(n\) counts field sources \(\varphi\), \(m\) counts density sources
\(J\), \(\ell\) counts fermion fields \(\psi\), and
\(|p|_1=\sum_i p_i\) counts derivatives on the fermion fields.  The associated
coordinate is a translation-invariant kernel
\[
  H_\alpha(x_1,\ldots,x_n;y_1,\ldots,y_m;z_1,\ldots,z_\ell)
\]
with the internal antisymmetries imposed by the Grassmann variables and with
finite pinned \(L^1\) norm, obtained by integrating the absolute value over all
but one spacetime variable and then taking the supremum over internal indices.
The fixed point is therefore a sequence \(H=\{H_\alpha\}_\alpha\), not a finite
polynomial in local fields.

Let the fermion scaling dimension be
\[
  [\psi]=\frac d4-\frac{\epsilon}{2},
\]
and let the source fields \(\varphi\) and \(J\) scale with dimensions
\(d-\Delta_1\) and \(d-\Delta_2\).  Equivalently, \(\Delta_1\) and
\(\Delta_2\) are the dimensions of the scaling operators whose response
functions are generated by differentiating with respect to \(\varphi\) and
\(J\).  If the scale factor is \(\gamma\), the dilatation of a pinned
\(L^1\) kernel has exponent
\begin{equation}
  D_{\rm sc}(\alpha)
  =
  d
  -n(d-\Delta_1)
  -m(d-\Delta_2)
  -\ell[\psi]
  -|p|_1 .
  \label{eq:fermionic-rg-kernel-scaling-dimension}
\end{equation}
Thus \(\|D H_\alpha\|_1=\gamma^{D_{\rm sc}(\alpha)}\|H_\alpha\|_1\) in the
kernel norm before the additional Gevrey or stretched-exponential weights are
applied.  With \(\Delta_1=[\psi]+O(\epsilon)\) and
\(\Delta_2=2[\psi]+O(\epsilon)\), the exceptional local block separated by
the trimming chart consists of the mass-type bilinear \((0,0,2,0)\), the
quartic \((0,0,4,0)\), the source-field coordinate \((1,0,1,0)\), the
density-source coordinate \((0,1,2,0)\), and the first-derivative bilinears
\((0,0,2,p)\) with \(|p|_1=1\).  This list is not being introduced as a
finite truncation of the theory: \((0,1,2,0)\) may be weakly irrelevant
depending on the sign of \(\Delta_2-2[\psi]\), and the first-derivative
bilinears are irrelevant in \(d=1\) for small \(\epsilon\).  They are kept in
the local block because their scaling exponents are close enough to zero, or
because a uniform trimmed coordinate system for \(d=1,2,3\) makes the
localization/interpolation estimates stable.  For example,
\begin{equation}
  D_{\rm sc}(0,0,4,0)=d-4[\psi]=2\epsilon,
  \qquad
  D_{\rm sc}(0,0,2,0)=d-2[\psi]=\frac d2+\epsilon .
  \label{eq:fermionic-rg-quartic-mass-dimensions}
\end{equation}
The quartic coordinate is weakly relevant for \(\epsilon>0\), which is why a
small non-Gaussian fixed point can be searched for near the Gaussian one.

The invariant space is obtained by a trimming operation.  The kernels in the
finite exceptional local list are represented by local delta-functions with
constant coefficients.  The nonlocal part of those same kernels is subtracted
and rewritten in derivative coordinates, so that it enters the irrelevant
Banach component.  All remaining coordinates live in a kernel space whose norm
controls pinned \(L^1\) size, internal index sums, scale weights, and
short-range decay.  Schematically, one writes
\[
  H=(g,K),
  \qquad
  g\in\mathbb C^r,\quad K\in\mathcal K,
\]
where \(g\) is the finite vector of local trimmed coordinates and
\(\mathcal K\) is the Banach space of irrelevant nonlocal kernels.  One RG step
has the form
\begin{equation}
  (g,K)\longmapsto
  \bigl(A_{\rm loc}g+B_{\rm loc}(g,K),\,
        A_{\rm irr}K+B_{\rm irr}(g,K)\bigr),
  \label{eq:fermionic-rg-trimmed-coordinate-map}
\end{equation}
after fluctuation integration, trimming, and rescaling.  The theorem's
analytic burden is to prove this map on a ball in
\(\mathbb C^r\oplus\mathcal K\), with estimates of the form
\begin{equation}
  \|A_{\rm irr}\|\leq q<1,
  \qquad
  \|B_{\rm irr}(g,K)-B_{\rm irr}(g',K')\|
  \leq
  c\,(|g|+|g'|+\|K\|+\|K'\|)\,
  (|g-g'|+\|K-K'\|),
  \label{eq:fermionic-rg-irrelevant-contraction-form}
\end{equation}
and corresponding finite-dimensional equations for \(g\).  The fixed point is
the simultaneous solution of both the local equation and
\begin{equation}
  K=A_{\rm irr}K+B_{\rm irr}(g,K).
  \label{eq:fermionic-rg-irrelevant-tail-equation}
\end{equation}
Dropping \eqref{eq:fermionic-rg-irrelevant-tail-equation} would reduce the
argument to a projected beta-function calculation and would miss exactly the
part of the theorem that makes the construction nonperturbative.

The determinant expansion used in the fermionic construction is also part of
the mechanism.  The Grassmann Gaussian integration reorganizes the multiscale
series into determinants of covariance matrices, and Gram-type determinant
bounds give absolute convergence in the kernel norm.  This is why the
fermionic example presently reaches a level of rigor not available for the
ordinary short-range scalar critical point: the invariant kernel space and the
convergent determinant/tree expansion are proved together.  The monograph uses
this as a proof template, not as evidence that the scalar or gauge targets are
settled.

\begin{remark}[Irrelevant tails are part of the theorem]
\label{rem:irrelevant-tail-part-of-rg-theorem}
Let \(\mathcal B=\mathbb R^m\oplus\mathcal K\), where \(\mathcal K\) is a
Banach space of irrelevant kernels, and let \(P:\mathcal B\to\mathbb R^m\) be
the coordinate projection.  A solution of the projected equation
\[
  P(\mathcal R(g,0)-(g,0))=0
\]
does not define a fixed point of \(\mathcal R\) unless the irrelevant residual
\[
  Q(\mathcal R(g,0)-(g,0))\in\mathcal K
\]
is either zero or is absorbed by a controlled graph \(K=h(g)\) satisfying the
full fixed-point equation.
Write \(Q=1-P\).  If \(V=(g,0)\) solves only the projected equation, then
\[
  \mathcal R(V)-V=(0,Q(\mathcal R(V)-V)).
\]
Thus \(V\) is a fixed point if and only if the displayed irrelevant residual
vanishes.  More generally, a fixed point with the same local coordinates must
have the form \(V_h=(g,h(g))\) and satisfy
\[
  P(\mathcal R(V_h)-V_h)=0,\qquad
  Q(\mathcal R(V_h)-V_h)=0 .
\]
The second equation is the irrelevant-tail equation.  In a constructive RG
proof it is solved by contraction, graph transform, or Newton validation in
the norm of \(\mathcal K\); without such a bound, the projected equation is
only a coordinate calculation.
\end{remark}

\section{Hierarchical Scalar Fixed-Point Benchmark}

Scalar non-Gaussian fixed points provide the most familiar physical target
for Wilsonian RG, but the mathematical status depends strongly on the map
being studied.  The hierarchical scalar models form a class in which the
block-spin map is an explicitly defined nonlinear integral operator.  They
are not the short-range nearest-neighbor lattice \(\phi^4\) models, and a
theorem for the hierarchical map is not, by itself, a theorem for the
ordinary three-dimensional Ising or \(\phi^4\) fixed point.  They are
nevertheless an essential benchmark because the fixed point, unstable
directions, and irrelevant remainder can be treated in a genuine
infinite-dimensional RG problem rather than in a finite beta-function
truncation.

\begin{definition}[Hierarchical scalar RG system]
\label{def:hierarchical-scalar-rg-system}
A one-component hierarchical scalar RG system consists of:
\begin{enumerate}
  \item a scale factor \(L>1\), a dimension parameter \(D>0\), and
  \(b=L^D\);
  \item a field scaling exponent \(\Delta>0\), a number
  \(a=L^{-\Delta}\), and a fluctuation variance \(\gamma>0\);
  \item a Banach space \(\mathcal B_{\rm hier}\) of even real-analytic
  functions \(V:\mathbb R\to\mathbb R\), modulo additive constants, with a
  norm controlling a complex strip or disk and large-field growth;
  \item the nonlinear map \(\mathcal R_{\rm hier}\) defined, whenever the
  integral is finite and analytic in the chosen domain, by
  \begin{equation}
    e^{-(\mathcal R_{\rm hier}V)(\Phi)}
    =
    Z_V^{-1}
    \int_{\mathbb R}
      \exp\!\left[-b\,V(a\Phi+\zeta)\right]\,
      \frac{e^{-\zeta^2/(2\gamma)}}{\sqrt{2\pi\gamma}}\,\dd\zeta ,
    \label{eq:hierarchical-scalar-rg-map}
  \end{equation}
  where \(Z_V\) is fixed by the normalization
  \((\mathcal R_{\rm hier}V)(0)=0\).
\end{enumerate}
This is a mathematical RG system.  The hierarchical covariance and the
normalization convention are part of the definition; they are not inferred
from a continuum path integral after the fact.
\end{definition}

The linearization at the Gaussian point already shows the local-coordinate
arithmetic that underlies the usual relevance labels.  The following
calculation is exact for the hierarchical Gaussian kernel and does not use a
formal momentum shell.

\paragraph{Gaussian linearization in Wick coordinates.}
Let \(0<a<1\), \(\gamma>0\), and
\[
  c_\ast=\frac{\gamma}{1-a^2}.
\]
For \(n\geq0\), define the Wick polynomial
\(:x^n:_{c_\ast}\) by
\begin{equation}
  \exp\!\left(tx-\frac{c_\ast t^2}{2}\right)
  =
  \sum_{n=0}^\infty \frac{t^n}{n!}\,:x^n:_{c_\ast}.
  \label{eq:hierarchical-wick-generating-function}
\end{equation}
The derivative of \(\mathcal R_{\rm hier}\) at \(V=0\), before quotienting
the additive constant, acts on the nonconstant Wick coordinates by
\begin{equation}
  D\mathcal R_{\rm hier}|_0\bigl(:\Phi^n:_{c_\ast}\bigr)
  =
  b\,a^n :\Phi^n:_{c_\ast},
  \qquad n\geq1 .
  \label{eq:hierarchical-wick-eigenvalue}
\end{equation}
In the canonical scalar bookkeeping
\[
  \Delta=\frac{D-2}{2},
  \qquad b=L^D,
\]
the even local coordinate \(:\Phi^{2r}:\) has engineering eigenvalue
\begin{equation}
  \lambda_{2r}=L^{y_{2r}},
  \qquad
  y_{2r}=D-r(D-2).
  \label{eq:hierarchical-scalar-engineering-exponent}
\end{equation}
Thus \(y_2=2\), \(y_4=4-D\), and \(y_6=6-2D\).
For the first variation of \eqref{eq:hierarchical-scalar-rg-map}, additive
normalization affects only the constant coordinate.  Hence for a
nonconstant test direction \(f\),
\begin{equation}
  D\mathcal R_{\rm hier}|_0 f
  =
  b\int_{\mathbb R} f(a\Phi+\zeta)\,\dd\mu_\gamma(\zeta)
  \quad\text{mod constants},
  \label{eq:hierarchical-linearized-map}
\end{equation}
where \(\dd\mu_\gamma\) is the centered Gaussian measure of variance
\(\gamma\).  Apply this to the generating function
\eqref{eq:hierarchical-wick-generating-function}.  Since
\(c_\ast-\gamma=a^2c_\ast\),
\[
\begin{aligned}
  \int
  \exp\!\left(t(a\Phi+\zeta)-\frac{c_\ast t^2}{2}\right)
  \dd\mu_\gamma(\zeta)
  &=
  \exp\!\left(ta\Phi-\frac{(c_\ast-\gamma)t^2}{2}\right)        \\
  &=
  \exp\!\left(ta\Phi-\frac{a^2c_\ast t^2}{2}\right)             \\
  &=
  \sum_{n=0}^\infty
    \frac{t^n}{n!}\,a^n:\Phi^n:_{c_\ast}.
\end{aligned}
\]
Comparing coefficients gives the factor \(a^n\); the prefactor \(b\) in
\eqref{eq:hierarchical-linearized-map} gives
\eqref{eq:hierarchical-wick-eigenvalue}.  Substituting
\(a=L^{-(D-2)/2}\) and \(n=2r\) gives
\[
  b\,a^{2r}
  =
  L^D L^{-r(D-2)}
  =
  L^{D-r(D-2)} ,
\]
which is \eqref{eq:hierarchical-scalar-engineering-exponent}.

\begin{quotedtheorem}[Hierarchical scalar non-Gaussian fixed point]
\label{qthm:hierarchical-scalar-nongaussian-fixed-point}
For the Dyson--Wilson hierarchical scalar RG maps in the
Bleher--Sinai and Collet--Eckmann class, and for the closely related
\(\phi^4_{3,\epsilon}\) RG construction of
Brydges--Mitter--Scoppola,\footnote{See P.~M.~Bleher and Ya.~G.~Sinai,
``Investigation of the critical point in models of the type of Dyson's
hierarchical models,'' \emph{Commun. Math. Phys.} \textbf{33} (1973);
P.~Collet and J.-P.~Eckmann, \emph{A Renormalization Group Analysis of the
Hierarchical Model in Statistical Mechanics}, Lecture Notes in Physics
\textbf{74}, Springer (1978); and D.~C.~Brydges, P.~K.~Mitter, and
B.~Scoppola, ``Critical \((\Phi^4)_{3,\epsilon}\),'' arXiv:hep-th/0206040.}
there are parameter regimes in which a
non-Gaussian fixed point exists as an element of a Banach space of analytic
or polymer interactions.  In these regimes the theorem supplies an RG map, a
domain and norm, a fixed interaction \(V_\ast\), a finite-dimensional
unstable subspace, contraction or graph estimates for the stable directions,
and scaling estimates for the corresponding hierarchical or
\(\phi^4_{3,\epsilon}\) observable data.
\end{quotedtheorem}

\paragraph{Status of the quoted theorem.}
The theorem is quoted as a benchmark, not as an imported proof for the
ordinary short-range three-dimensional scalar fixed point.  Its mathematical
content is that the non-Gaussian fixed point is an element of the specified
infinite-dimensional RG domain and that the irrelevant coordinates are
controlled by estimates in that domain.  The hierarchical proofs exploit the
recursive structure of the covariance; the
\(\phi^4_{3,\epsilon}\) construction uses a multiscale decomposition and
polymer estimates adapted to the fractional covariance.  In both cases the
result is stronger than locating a zero of a finite beta function and weaker
than a construction of the short-range \(\phi^4_3\) Wilson--Fisher CFT.
That last comparison would require a theorem relating the short-range
block-spin map and reconstruction data to one of these fixed-point systems.

The companion calculation
\[
  \texttt{calculation-checks/rg\_hierarchical\_scalar\_checks.py}
\]
checks \eqref{eq:hierarchical-wick-eigenvalue} and
\eqref{eq:hierarchical-scalar-engineering-exponent} with exact rational
arithmetic for the even Wick coordinates used above.

\section{Current Rigorous Fixed-Point Benchmarks}

The preceding definition deliberately separates what has been proved from
what is inferred from formal or numerical RG.  Recent rigorous work supplies
benchmarks for both the constructive-field-theory and tensor-network
versions of the problem.

The hierarchical scalar results just described are the cleanest scalar
examples of theorem-level non-Gaussian Wilsonian fixed points.  Their role
in this chapter is deliberately calibrated: they prove that scalar
non-Gaussian fixed points can be constructed in rigorous RG systems with
irrelevant tails, but they do not settle the ordinary short-range
three-dimensional scalar fixed point.

In the long-range fermionic \(\psi^4_d\) models analyzed by
Giuliani, Mastropietro, Rychkov, and collaborators,\footnote{See
A.~Giuliani, V.~Mastropietro, and S.~Rychkov, ``Gentle introduction to
rigorous Renormalization Group: a worked fermionic example,''
JHEP \textbf{01} (2021) 026, arXiv:2008.04361; and
A.~Giuliani, V.~Mastropietro, S.~Rychkov, and G.~Scola,
``Non-trivial fixed point of a \(\psi^4_d\) fermionic theory, II. Anomalous
exponent and scaling operators,'' arXiv:2404.14904.}
the Banach space contains all short-ranged irrelevant kernels, not only a
finite list of couplings.  The fixed interaction is constructed by a
convergent RG scheme with a small parameter controlling the relevance of the
quartic interaction.  The important lesson for this monograph is structural:
the fixed point is an element of an infinite-dimensional space, the irrelevant
tail is part of the theorem rather than a truncation error, and scaling
operators are constructed as response functions with stated remainder
estimates.

In the rigorous tensor-RG work of Kennedy and Rychkov,\footnote{See
T.~Kennedy and S.~Rychkov, ``Tensor RG approach to high-temperature fixed
point,'' J. Stat. Phys. \textbf{187} (2022) 33, arXiv:2107.11464; and
``Tensor Renormalization Group at Low Temperatures: Discontinuity Fixed
Point,'' arXiv:2210.06669.}
the map acts on infinite-dimensional tensors.  The theorem is not a statement
about the naive contraction of four tensors, which fails to be contractive
near the high-temperature fixed point; it is a statement about a modified RG
map with a disentangling step and a Hilbert--Schmidt neighborhood on which
contraction estimates are proved.  The low-temperature analysis gives stable
ordered fixed points and a discontinuity fixed point for first-order
transition behavior.  These examples illustrate why the RG map itself is part
of the theorem: changing the coarse-graining prescription changes the
mathematical assertion being made.

Functional RG equations such as the Wetterich equation are exact identities
at finite regulator, as derived in
Chapter~\ref{ch:vol2-wilsonian-effective-actions}.  A finite-dimensional
truncation of such an equation can be an excellent computational scheme, but
the theorem-level fixed-point claim requires an error estimate comparing the
truncated subspace with the full Banach space.  The same distinction applies
to perturbative epsilon expansions: the formal fixed point organizes local
coordinates and critical exponents order by order, while a nonperturbative
fixed-point theorem additionally proves existence, stability, and
reconstruction.

\section{Fixed-Point Claim Content}
\label{sec:rg-fixed-point-claim-content}

The preceding benchmarks should be compared through the objects they actually
construct.  A fixed-point statement in this chapter is compared through four
questions:
\begin{enumerate}
  \item what space the RG map acts on, such as local interactions, polymer
  activities, Grassmann kernels, tensors, lattice measures, or BV actions;
  \item what estimates prove that the map is defined on its domain, is
  differentiable or analytic there, has the claimed fixed point, and has the
  stated stable/unstable decomposition;
  \item which source extensions, contact-term subspaces, and reconstruction
  estimates turn the fixed point into distributions or observable algebras;
  \item which comparison map identifies the constructed RG system with the
  physical short-range, gauge, or continuum QFT target.
\end{enumerate}
The strength of a result is the first question in this list for which the
necessary objects and estimates are missing.  This convention prevents an
auxiliary fixed point from being silently promoted to a theorem about a
different QFT.

\paragraph{Hierarchical scalar systems.}
For the hierarchical scalar maps of
Definition~\ref{def:hierarchical-scalar-rg-system}, the RG object and map
theorem are part of the rigorous theorem: the nonlinear integral map is
defined on a Banach space of analytic or polymer interactions, the
non-Gaussian fixed point is constructed in that space, and stable/unstable
directions are controlled by the corresponding estimates.  Observable
statements are hierarchical observable statements unless accompanied by a
short-range reconstruction estimates and a transfer map of the form
Definitions~\ref{def:short-range-scalar-block-spin-structure} and
\ref{def:auxiliary-to-short-range-rg-transfer-estimates}.  The target
identification with the ordinary nearest-neighbor or finite-range
three-dimensional scalar critical point is therefore an additional problem,
requiring its own comparison map and reconstruction estimates beyond shared
coordinate names such as scalar, relevant, or quartic.

\paragraph{Long-range fermionic models.}
For the long-range fermionic \(\psi^4_d\) systems, the rigorous result
constructs a fixed interaction in a Banach space of Grassmann kernels and
constructs selected response functions with anomalous scaling and controlled
remainders.  The source layer is part of the theorem for the field and
density response functions that are actually constructed.  The target is the
long-range fermionic model with the specified fractional kinetic term and
fixed ultraviolet cutoff interpretation.  Using this result as evidence for
ordinary short-range scalar critical behavior requires a separately stated
auxiliary-to-target comparison; it is not an identification of universality
classes by analogy.

\paragraph{Tensor RG fixed points.}
For tensor RG, the RG object is an infinite-dimensional tensor in a declared
Hilbert--Schmidt or Banach topology.  The theorem is a contraction statement
for a modified tensor map, including its disentangling or truncation rule,
on a specified neighborhood of the high-temperature, low-temperature, or
discontinuity fixed point.  The reconstructed object is initially a lattice
statistical model or tensor-network partition function.  A continuum-QFT
claim requires a further scaling-window reconstruction and comparison to
Schwinger functions, local algebras, or another QFT observable topology.
Thus tensor RG supplies theorem-level fixed points of tensor maps and a
valuable laboratory for RG proof technology; it becomes a continuum QFT
fixed-point theorem only after the reconstruction layer is supplied.

\paragraph{Ordinary short-range scalar targets.}
For a short-range three-dimensional scalar Wilson--Fisher target, the RG
object must be the block-spin/polymer structure of
Definition~\ref{def:short-range-scalar-block-spin-structure}, or a proven
equivalent structure connected to it by
Definition~\ref{def:comparison-of-rg-descriptions}.  The theorem would have
to construct the stable critical trajectory or fixed point in that structure,
including the irrelevant polymer tail, source extensions, contact terms, and
distributional reconstruction estimates.  The massive constructive
\(\phi^4_3\) theorems and the critical hierarchical or long-range
fixed-point theorems supply important pieces of the landscape, but the
ordinary short-range critical reconstruction theorem is a separate
target-identification problem.  Perturbative epsilon expansions and
finite-dimensional functional-RG truncations are then used as coordinate and
numerical evidence for this target, not as substitutes for the map theorem
and reconstruction estimates.

\paragraph{Gauge targets.}
For gauge theories, the RG object is not a list of gauge-fixed coefficients.
The object is either a gauge-invariant lattice measure and observable system,
or a regularized BV action and observable complex satisfying the quantum
master equation and its RG pushforward.  The source layer must use
gauge-invariant local observables, Wilson-line or Wilson-loop observables
with their geometric data, or BV cohomology classes.  A gauge-theory
fixed-point or continuum-limit theorem therefore includes preservation of
the gauge structure under blocking, reconstruction of physical observables,
and comparison of line or extended sectors when those sectors are part of
the claim.  Beta functions in a gauge-fixed coordinate chart are useful
local coordinates only after this gauge-compatible RG structure has been
declared.

This comparison also fixes how this monograph treats the phrase ``state of the
art.''  The current rigorous fixed-point frontier contains theorem-level
auxiliary and model-specific systems with impressive control of irrelevant
coordinates.  It does not yet contain a general theorem turning every
physically expected Wilsonian fixed point into a local QFT, nor a theorem
constructing the ordinary three-dimensional Ising CFT from a short-range
block-spin map with all source and reconstruction estimates.  The monograph uses
the proved systems as proof templates and laboratories, while keeping the
physical target-identification problems explicit.

\section{Gauge-Compatible RG And Reconstruction Estimates}

A Wilsonian RG theorem for a scalar probability measure can be formulated as
a pushforward of measures, followed by a localization step and a rescaling.
Gauge theories require richer structure, because the variables used in a
calculation contain redundancy.  The theorem-level object is therefore an RG
map on gauge-invariant lattice measures and observables, or an RG map in the BV category whose
effective actions and observable complexes satisfy the quantum master
equation at every scale.  The BV framework used here is developed in
Chapter~\ref{ch:bv-master-formalism}; the lattice Yang--Mills and reflection
positivity inputs are developed in
Chapters~\ref{ch:lattice-yang-mills-nonperturbative-definition},
\ref{ch:lattice-reflection-positivity}, and
\ref{ch:constructive-wilson-lattice-gauge-theory}.

\begin{definition}[Gauge-compatible Wilsonian RG structure]
\label{def:gauge-compatible-wilsonian-rg-structure}
Fix a spacetime dimension \(D\), a scale factor \(L>1\), and a regulator
sequence indexed by \(k\in\mathbb Z_{\geq0}\).  A gauge-compatible Wilsonian
RG structure is one of the following two forms.
\begin{enumerate}
  \item In the lattice form, one has finite cell complexes
  \(\Lambda_k\), a compact Lie group \(G_k\), configuration spaces
  \(\mathcal A_k=G_k^{E(\Lambda_k)}\) of link variables, gauge groups
  \(\mathcal G_k=G_k^{V(\Lambda_k)}\) acting on \(\mathcal A_k\), and finite
  measures \(\dd\mu_k\) on \(\mathcal A_k\).  The observables are functions or
  distributions on \(\mathcal A_k\) invariant under \(\mathcal G_k\), with a
  declared norm controlling Wilson loops, local plaquette fields, and any
  matter insertions with their Wilson-line connectors.  The RG step is a
  gauge-equivariant blocking kernel
  \[
    B_{k,k-1}(A_{k-1},A_k)\,\dd\mu_k(A_k)
  \]
  whose pushforward gives a measure \(\dd\mu_{k-1}\) on
  \(\mathcal A_{k-1}\), followed by a localization map into a Banach chart of
  gauge-invariant polymer activities.  Reflection positivity, when claimed,
  is part of the structure and is expressed as positivity of the blocked
  quadratic form on the positive-time observable algebra.
  \item In the BV form, one has finite-regulator graded vector spaces
  \(\mathcal F_k\) of fields, ghosts, antifields, and possible sources; odd
  symplectic forms of degree \(-1\); BV Laplacians \(\Delta_k\) associated
  with the chosen finite-dimensional Berezin-Lebesgue density; and actions
  \(S_k\) satisfying the quantum master equation
  \begin{equation}
    \frac12(S_k,S_k)_k-\hbar\,\Delta_k S_k=0 ,
    \label{eq:gauge-compatible-qme-at-scale}
  \end{equation}
  or equivalently \(\Delta_k e^{-S_k/\hbar}=0\).  The RG step is BV
  pushforward along a fluctuation subspace with gauge-fixing Lagrangian
  \(\mathcal L_k\):
  \begin{equation}
    e^{-S_{k-1}(\phi_{k-1})/\hbar}
    =
    \int_{\mathcal L_k}
      e^{-S_k(\phi_{k-1}+\xi_k)/\hbar}\,\dd\xi_k ,
    \label{eq:bv-rg-pushforward}
  \end{equation}
  together with a localization map into a Banach space of local BV
  functionals.  The induced observable map
  \(P_{k,k-1}\) sends \(Q_{S_k}\)-closed observables to
  \(Q_{S_{k-1}}\)-closed observables, where
  \(Q_{S_k}=(S_k,\cdot)_k-\hbar\Delta_k\).
\end{enumerate}
In either form, a fixed point or continuum limit also includes reconstruction
maps for gauge-invariant local and extended observables.  A flow of finitely
many gauge-fixed coefficients is a coordinate shadow of this structure; the
theorem-level statement is the lattice or BV structure together with its
norm estimates.
\end{definition}

The definition deliberately includes Wilson-line and loop observables.  A
colored field variable in a gauge-fixed calculation is useful for local
coordinate work, but a nonperturbative observable statement is formulated in
terms of gauge-invariant insertions or BV cohomology classes.  Thus a
charged-particle or partonic reconstruction theorem must specify the line,
asymptotic dressing, or cohomology class used to define the observable before
one speaks of an RG limit.

\begin{construction}[Gauge-equivariant path blocking at finite regulator]
\label{constr:gauge-equivariant-path-blocking-rg}
The lattice half of Definition~\ref{def:gauge-compatible-wilsonian-rg-structure}
has a minimal exact example.  Let \(\Gamma_f\) and \(\Gamma_c\) be finite
oriented graphs, interpreted as fine and coarse lattices.  Let
\(i:V(\Gamma_c)\hookrightarrow V(\Gamma_f)\) embed coarse vertices into fine
vertices.  For every oriented coarse edge \(e:v\to w\), choose an oriented
fine path
\[
  P_e=(\ell_1,\ldots,\ell_m),
  \qquad
  s(\ell_1)=i(v),\quad t(\ell_m)=i(w),
\]
and require \(P_{\bar e}=\overline{P_e}\), the same path with reversed
orientation.  A fine gauge field assigns \(U_\ell\in G\) to oriented fine
edges with \(U_{\bar\ell}=U_\ell^{-1}\).  Define the blocked coarse link by
\begin{equation}
  B(A)_e
  =
  U_{\ell_1}U_{\ell_2}\cdots U_{\ell_m}.
  \label{eq:gauge-path-blocked-link}
\end{equation}
For a fine gauge transformation \(g\in G^{V(\Gamma_f)}\),
\[
  U_\ell^g=g_{s(\ell)}U_\ell g_{t(\ell)}^{-1}.
\]
The internal endpoint factors along \(P_e\) cancel, so
\begin{equation}
  B(A^g)_e
  =
  g_{i(v)}\,B(A)_e\,g_{i(w)}^{-1}.
  \label{eq:gauge-path-blocking-equivariance}
\end{equation}
Thus \(B\) is equivariant from the fine gauge action to the coarse gauge
action obtained by restricting \(g\) to embedded coarse vertices.

Let \(\mu_f\) be any finite measure on \(G^{E(\Gamma_f)}\) invariant under
the fine gauge group, and let \(\mu_c=B_\ast\mu_f\).  For a coarse gauge
transformation \(h\in G^{V(\Gamma_c)}\), choose any extension
\(\tilde h\in G^{V(\Gamma_f)}\) with
\(\tilde h_{i(v)}=h_v\).  Equation
\eqref{eq:gauge-path-blocking-equivariance} gives
\[
  B(A^{\tilde h})=B(A)^h .
\]
Since \(\mu_f\) is invariant under \(A\mapsto A^{\tilde h}\), the
pushforward \(\mu_c\) is invariant under the coarse gauge group.  If
\(C=e_1\cdots e_r\) is a closed coarse loop and \(R\) is a finite-dimensional
unitary representation of \(G\), then
\begin{equation}
  W_R(C;B(A))
  =
  \frac{1}{\dim R}
  \operatorname{tr}_R
  \bigl(B(A)_{e_1}\cdots B(A)_{e_r}\bigr)
  \label{eq:gauge-path-blocked-wilson-loop}
\end{equation}
is the Wilson loop of the concatenated fine path and is gauge invariant.

This construction proves only the exact finite-regulator covariance and
observable pushforward.  It does not prove that the logarithm of the blocked
measure is a local Wilson action, nor that a continuum local QFT follows.
Those are precisely the localization, polymer-decay, reflection-positivity,
and reconstruction estimates required in the rest of
Definition~\ref{def:gauge-compatible-wilsonian-rg-structure}.
\end{construction}

The companion script
\texttt{calculation-checks/lattice\_gauge\_blocking\_checks.py} verifies the
finite group algebra behind
\eqref{eq:gauge-path-blocking-equivariance} and
\eqref{eq:gauge-path-blocked-wilson-loop} for an \(S_3\) plaquette example,
including invariance of the blocked pushforward weights under coarse gauge
transformations.

\section{Gauge-Blocking Locality, Positivity, And Reconstruction Estimates}
\label{sec:gauge-blocking-locality-positivity-reconstruction}

The path-blocking construction above is an exact gauge-covariance statement.
It is not yet a Wilsonian RG theorem for a continuum gauge theory.  A
continuum claim needs three further pieces of analysis: locality of the
blocked measure in a declared norm, reflection positivity or a stated
replacement for it, and reconstruction estimates for gauge-invariant
observables.  This section records the estimates required for a theorem-level
four-dimensional Yang--Mills statement.

Let \(\Gamma_k\) be a sequence of Euclidean lattice complexes with lattice
spacing \(a_k\to0\).  For a connected set \(X\) of coarse plaquettes, links,
or cells, let \(\operatorname{diam}_k(X)\) be its diameter in physical units
and let \(\operatorname{dist}_k(X,Y)\) be the physical distance between two
cell sets.  Let \(\operatorname{supp}F\) be the smallest connected cell set on
which a cylinder observable \(F\) depends.  The norm
\(\|\,\cdot\,\|_\infty\) is the supremum norm on the corresponding compact
link-configuration space.

\begin{definition}[Gauge-blocking continuum-control estimates]
\label{def:gauge-blocking-continuum-control-datum}
A continuum-control estimate package for a gauge-compatible blocking sequence consists
of the finite path-blocking data of
Construction~\ref{constr:gauge-equivariant-path-blocking-rg}, together with
the following estimates.
\begin{enumerate}
  \item A gauge-invariant polymer representation of the blocked measure at
  scale \(k\):
  \begin{equation}
    \frac{\dd\mu_k^{\rm blk}}{\dd\nu_k}(A)
    =
    Z_k^{-1}
    \exp\!\left(-\sum_{X\subset\Gamma_k}K_{k,X}(A_X)\right),
    \label{eq:gauge-blocked-polymer-representation}
  \end{equation}
  where \(\nu_k\) is a reflection-positive local reference measure,
  \(K_{k,X}\) is gauge invariant and depends only on links meeting \(X\),
  and the sum is over connected polymers \(X\).
  \item A quasi-locality estimate: for some \(\zeta>0\),
  \begin{equation}
    \|K_k\|_{\zeta}
    =
    \sup_{e\in E(\Gamma_k)}
    \sum_{X\ni e}
      e^{\zeta\operatorname{diam}_k(X)}
      \|K_{k,X}\|_\infty
    \leq \varepsilon_k ,
    \qquad
    \varepsilon_k\longrightarrow0
    \label{eq:gauge-blocking-polymer-locality-norm}
  \end{equation}
  after subtracting the finitely many local coordinates retained in the
  Wilsonian chart.  In addition, for every bounded gauge-invariant cylinder
  observable \(F\), the cluster expansion or polymer estimate supplies
  truncated expectations \(\langle F\rangle_{k,R}\), obtained by keeping only
  connected polymer clusters contained in the \(R\)-neighborhood of
  \(\operatorname{supp}F\), and constants \(C_F\) such that
  \begin{equation}
    \left|
      \langle F\rangle_k-\langle F\rangle_{k,R}
    \right|
    \leq
    C_F e^{-\zeta R}\|K_k\|_{\zeta}
    +\chi_{F,k}(R),
    \qquad
    \lim_{R\to\infty}\sup_k\chi_{F,k}(R)=0 .
    \label{eq:gauge-blocking-cluster-tail-estimate}
  \end{equation}
  This cluster-tail estimate is part of the continuum-control package, not a
  consequence of gauge covariance alone.
  \item Reflection-positivity control.  Let
  \(\mathcal A_{k,+}^{\rm inv}\) be the algebra of gauge-invariant cylinder
  observables supported in the positive-time half-lattice, and let \(\Theta\)
  be time reflection with orientation reversal on links crossing reflected
  edges.  For every finite list \(F_1,\ldots,F_m\in\mathcal A_{k,+}^{\rm inv}\)
  and coefficients \(c_i\in\C\),
  \begin{equation}
    \sum_{i,j=1}^{m}\bar c_i c_j
    \int \Theta F_i(A)\,F_j(A)\,\dd\mu_k^{\rm blk}(A)
    \geq -\rho_k(F,c),
    \label{eq:gauge-blocking-rp-defect}
  \end{equation}
  with \(\rho_k(F,c)\to0\) for each fixed finite family.  Exact reflection
  positivity is the special case \(\rho_k=0\).
  \item Gauge-invariant reconstruction maps for a declared class of
  observables \(\mathcal W\), such as finite Wilson-loop products, smeared
  small-loop field-strength representatives, or matter bilinears with their
  Wilson-line connectors.  For each \(O\in\mathcal W\), the reconstructed
  finite-scale observable \(\operatorname{Rec}_{O,k}\) takes values in a locally
  convex distribution or cylinder-observable topology and satisfies a
  bound of the form
  \begin{equation}
    p\!\left(\operatorname{Rec}_{O,k}-\operatorname{Rec}_{O,\ast}\right)
    \leq
    C_{O,p}\varepsilon_k+\eta_{O,p,k},
    \qquad
    \eta_{O,p,k}\longrightarrow0 ,
    \label{eq:gauge-blocking-observable-reconstruction}
  \end{equation}
  for every seminorm \(p\) used in the continuum statement.
\end{enumerate}
\end{definition}

The first item is not a harmless rewriting.  The logarithm of a blocked
gauge-invariant measure can contain arbitrarily large Wilson loops.  The
locality estimate \eqref{eq:gauge-blocking-polymer-locality-norm} is the
assertion that those large connected contributions are summable in a norm
strong enough to define a Wilsonian chart.  In gauge theory this estimate is
one of the main constructive burdens.

\paragraph{Consequences of the estimates.}
Once Definition~\ref{def:gauge-blocking-continuum-control-datum} is proved in
a model, every \(O\in\mathcal W\) has the declared continuum limit by
\eqref{eq:gauge-blocking-observable-reconstruction}.  Gauge invariance passes
to the limit because it holds at finite regulator by path-blocking
equivariance and because the reconstruction topology tests gauge-invariant
cylinder data.  The estimate
\eqref{eq:gauge-blocking-cluster-tail-estimate} is the quasi-locality
statement: distant connected polymer clusters have a uniformly vanishing
effect on each bounded local gauge-invariant observable after the retained
local Wilsonian coordinates have been removed.

Reflection positivity is also a closed-cone statement once the finite
quadratic forms have been identified.  For
\(F_1,\ldots,F_m\in\mathcal A_{k,+}^{\rm inv}\) and \(c_i\in\C\), write
\[
  Q_k(F,c)=
  \sum_{i,j}\bar c_i c_j
  \int \Theta F_i\,F_j\,\dd\mu_k^{\rm blk}.
\]
The reconstruction estimate gives convergence of the finitely many
correlators entering \(Q_k(F,c)\), while
\eqref{eq:gauge-blocking-rp-defect} gives
\(Q_k(F,c)\geq-\rho_k(F,c)\).  Taking \(k\to\infty\) gives the limiting
positive-time quadratic form \(Q_\ast(F,c)\geq0\).  This is why the estimate package must
include the reflection-positive observable algebra and the defect
\(\rho_k\); it is not enough to know that the blocked links transform
covariantly.

\paragraph{Gauge-invariant source windows.}
The source-window layer in a gauge RG theorem is part of the datum, because
the source variables select which continuum distributions are being
reconstructed.  In the lattice form of
Definition~\ref{def:gauge-compatible-wilsonian-rg-structure}, let
\[
  E=\operatorname{span}_{\C}\{O_1,\ldots,O_N\}
\]
be a finite-dimensional space of gauge-invariant cylinder observables on
\(\mathcal A_k\).  Typical basis elements are closed Wilson loops, products
of small-loop field-strength representatives, or matter bilinears whose
color indices are contracted by a declared Wilson-line connector.  The path,
representation, endpoint contraction, and smearing data are part of the
observable label.  For \(z=(z_1,\ldots,z_N)\) in a fixed polydisc \(P_\rho\),
define
\begin{equation}
  W_{k,E}^{\rm inv}(z)
  =
  \log
  \int_{\mathcal A_k}
    \exp\!\left(\sum_{a=1}^N z_a O_a(A)\right)
    \dd\mu_k(A),
  \label{eq:gauge-invariant-source-window}
\end{equation}
whenever the integral is nonzero and the logarithm branch has been fixed at
\(z=0\).  Since each \(O_a\) is invariant under \(\mathcal G_k\), the
source-deformed density in
\eqref{eq:gauge-invariant-source-window} is invariant pointwise and therefore
descends to the orbit space or, equivalently, to the invariant observable
algebra.  An admissible source for a colored transport operator includes the
endpoint color contractions, Wilson-line geometry, or asymptotic dressing
that makes the observable gauge invariant; the uncontracted open link belongs
to a gauge-fixed coordinate chart.

In the BV form the same statement is slightly sharper.  A finite source
window is a holomorphic family \(S_k(z)\) with \(S_k(0)=S_k\) satisfying the
scale-\(k\) quantum master equation and equipped with its composite-operator
source coordinates.  The first source derivatives
\(\partial_{z_a}S_k(0)\) are quantum observables, meaning
\(Q_{S_k}\partial_{z_a}S_k(0)=0\).  The higher source derivatives of
\(S_k(z)\) are the contact and composite-operator coordinates needed for the
source family itself to remain in the BV chart.  Thus the distinction between
regularizing the path-integral measure and regularizing composite
operators is built into the RG source datum rather than added after
correlators have been computed.

The gauge source window is useful only if it obeys the same analytic
estimates as the scalar source windows above, now in the invariant or BV
cohomological chart.  A theorem-facing estimate has the form
\begin{equation}
  W_{k,E}^{\rm inv}(z)
  =
  C_{k,E}^{\rm inv}(z)
  +\mathfrak W_E^{\rm inv}
     \!\left(g_{k,E}^{\rm inv}(z),K_{k,E}^{\rm inv}(z)\right)
  +r_{k,E}^{\rm inv}(z),
  \label{eq:gauge-invariant-source-chart}
\end{equation}
on a fixed polydisc, with all three terms defined in the gauge-invariant
lattice chart or in a BV chart satisfying the quantum master equation.  If
the right side lies in a common Lipschitz neighborhood of the limiting
source chart, then
\begin{equation}
  \sup_{z\in P_\rho}
  \left|
    W_{k,E}^{\rm inv}(z)-W_{\ast,E}^{\rm inv}(z)
  \right|
  \leq
  \delta_{C,E}^{\rm inv}
  +L_E^{\rm inv}
    \bigl(
      \delta_{g,E}^{\rm inv}
      +\delta_{K,E}^{\rm inv}
    \bigr)
  +\delta_{r,E}^{\rm inv}.
  \label{eq:gauge-invariant-source-window-bound}
\end{equation}
The Cauchy estimates then extract connected correlators of the gauge
observables \(O_a\).  Equation
\eqref{eq:gauge-invariant-source-window-bound} is the place where a
finite-regulator gauge computation becomes a reconstruction statement.  A
finite list of Wilson-loop expectation values, or a beta function in a
gauge-fixed coordinate chart, becomes reconstruction input only after the
source radius, local coordinates, polymer tail, contact/source coordinates,
and thermodynamic boundary tails have been controlled.

The companion script
\texttt{calculation-checks/lattice\_gauge\_blocking\_checks.py} includes the
finite \(S_3\) check behind this paragraph: class functions of the blocked
closed loop give source polynomials constant on coarse gauge orbits, while a
source attached to a single open link fails the same invariance test.

\begin{warning}[Finite gauge blocking is not a continuum Yang--Mills proof]
\label{warn:finite-gauge-blocking-not-continuum-proof}
Construction~\ref{constr:gauge-equivariant-path-blocking-rg} and the finite
\(S_3\) companion calculation prove exact gauge covariance and exact
pushforward invariance at finite regulator.  They do not by themselves prove
the quasi-local polymer bound
\eqref{eq:gauge-blocking-polymer-locality-norm}, reflection positivity
\eqref{eq:gauge-blocking-rp-defect}, or reconstruction
\eqref{eq:gauge-blocking-observable-reconstruction}.  A theorem that a
gauge-blocked Wilsonian trajectory defines a continuum local QFT must supply
those estimates in a specific regulator class and observable topology.
\end{warning}

\begin{definition}[RG reconstruction estimate]
\label{def:rg-reconstruction-estimate}
Let \(\mathcal S_n=\mathcal S((\mathbb R^D)^n)\) be the Schwartz space and
let \(\mathcal S_n'\) be its continuous dual.  For a bounded set
\(B\subset\mathcal S_n\), write
\[
  p_B(T)=\sup_{f\in B}|\langle T,f\rangle|,
  \qquad T\in\mathcal S_n' .
\]
Let \(\mathcal R:\mathcal U\to\mathcal B\) be an RG map with fixed point
\(V_\ast\).  An \(n\)-point reconstruction estimate consists of distributions
\(\mathcal S_{n,\ast}\in\mathcal S_n'\), maps
\begin{equation}
  \operatorname{Rec}_{n,k}:\mathcal U\longrightarrow\mathcal S_n',
  \qquad k\in\mathbb Z_{\geq0},
  \label{eq:rg-reconstruction-map}
  \end{equation}
and constants \(C_B<\infty\), \(0<\theta<1\), and
\(\varepsilon_{B,k}\to0\), such that every tuned initial point \(V\) in the
declared stable manifold satisfies
\begin{equation}
  p_B\!\left(
    \operatorname{Rec}_{n,k}(V)-\mathcal S_{n,\ast}
  \right)
  \leq
  C_B\,\theta^k\,\|V-V_\ast\|_{\mathcal B}
  +\varepsilon_{B,k}
  \label{eq:rg-reconstruction-estimate}
\end{equation}
for all bounded \(B\subset\mathcal S_n\).  The maps
\(\operatorname{Rec}_{n,k}\) include the lattice spacing, coordinate
rescaling, field-strength normalization, source normalization, and possible
Wilson-line or BV-observable representative.
\end{definition}

\begin{proposition}[Reconstruction estimates produce distributional limits]
\label{prop:rg-reconstruction-estimates-give-limits}
Assume Definition~\ref{def:rg-reconstruction-estimate} for every
\(n\geq1\) and every bounded set \(B\subset\mathcal S_n\).  Then the
renormalized \(n\)-point functions
\(\operatorname{Rec}_{n,k}(V)\) converge to
\(\mathcal S_{n,\ast}\) in the strong dual topology of
\(\mathcal S_n'\).  If, at each finite regulator, the reconstructed
Schwinger functions are symmetric, Euclidean covariant, and reflection
positive, and if these properties are expressed by continuous linear or
closed positive-semidefinite inequalities in the seminorms \(p_B\), then the
limits \(\mathcal S_{n,\ast}\) satisfy the same properties.
\end{proposition}

\begin{proof}
Strong convergence in \(\mathcal S_n'\) means convergence with respect to the
seminorms \(p_B\) as \(B\) ranges over bounded subsets of \(\mathcal S_n\).
The estimate \eqref{eq:rg-reconstruction-estimate} gives
\[
  p_B(\operatorname{Rec}_{n,k}(V)-\mathcal S_{n,\ast})\longrightarrow0
\]
because \(\theta^k\to0\) and \(\varepsilon_{B,k}\to0\).  This proves
distributional convergence in the stated topology.

For symmetry and Euclidean covariance, let \(U\) denote a permutation or
Euclidean transformation acting continuously on \(\mathcal S_n\).  If the
finite-regulator distributions satisfy
\(\langle \operatorname{Rec}_{n,k}(V),f\rangle
=\langle \operatorname{Rec}_{n,k}(V),Uf\rangle\), then passing to the limit
in the two scalar sequences gives the same identity for
\(\mathcal S_{n,\ast}\).

For reflection positivity, take a finite family of positive-time test
functions \(f_i\) and coefficients \(c_i\).  At finite regulator the
corresponding quadratic expression has the form
\begin{equation}
  Q_k(c,f)=
  \sum_{i,j}\bar c_i c_j\,
  \bigl\langle
    \operatorname{Rec}_{m_i+m_j,k}(V),
    \Theta f_i\otimes f_j
  \bigr\rangle
  \geq0 ,
  \label{eq:rg-reflection-positivity-closed-limit}
\end{equation}
where \(\Theta\) is Euclidean time reflection.  Each term converges by the
already proved distributional convergence, hence \(Q_k(c,f)\to Q_\ast(c,f)\).
The cone \([0,\infty)\subset\mathbb R\) is closed, so \(Q_\ast(c,f)\geq0\).
Thus the limiting Schwinger functions inherit reflection positivity in the
declared test-function domain.  The same closed-cone argument applies to any
positivity statement explicitly formulated as a finite positive-semidefinite
quadratic form continuous in the seminorms \(p_B\).
\end{proof}

\section{Ordinary Short-Range Scalar Reconstruction}
\label{sec:ordinary-short-range-scalar-rg-reconstruction}

The scalar benchmark in the previous sections is hierarchical.  An ordinary
short-range scalar lattice model has a different microscopic object: fields
on \(a_k\mathbb Z^D\), finite-range interactions, and block-spin maps whose
range is geometric in the lattice.  The purpose of this section is to record
the reconstruction structure for that ordinary setting.  It is the structure a
proof of a short-range Wilson--Fisher fixed point would have to supply; it is not
replaced by a hierarchical or tensor-network theorem unless a comparison
map with reconstruction estimates is also proved.

\begin{definition}[Short-range scalar block-spin structure]
\label{def:short-range-scalar-block-spin-structure}
Fix an integer scale factor \(L\geq2\), spacetime dimension \(D\), and
lattice spacings \(a_k=L^{-k}a_0\).  A short-range scalar block-spin structure
consists of the following objects.
\begin{enumerate}
  \item Periodic lattice regions
  \(\Lambda_k\subset a_k\mathbb Z^D\), scalar configuration spaces
  \(\mathbb R^{\Lambda_k}\), and finite Gibbs measures
  \begin{equation}
    \dd\mu_k(\phi)
    =
    Z_k^{-1}
    \exp\!\left(
      -\sum_{\substack{X\subset\Lambda_k\\
        \operatorname{diam}(X)\leq R a_k}}
      U_{k,X}(\phi_X)
    \right)
    \prod_{x\in\Lambda_k}\dd\phi_x .
    \label{eq:short-range-scalar-gibbs-measure}
  \end{equation}
  Here \(R<\infty\) is fixed before the continuum limit, and the potentials
  are assumed measurable with enough coercivity to make all polynomial
  moments used below finite.
  \item A block kernel \(q:\{0,\ldots,L-1\}^D\to\mathbb R\) satisfying
  \begin{equation}
    \sum_{r\in\{0,\ldots,L-1\}^D}q(r)=1 ,
    \label{eq:block-kernel-normalization}
  \end{equation}
  and a field scaling exponent \(\Delta_\phi\).  The blocked field on
  \(\Lambda_{k-1}\) is
  \begin{equation}
    (B_k\phi)(y)
    =
    L^{\Delta_\phi}
    \sum_{r\in\{0,\ldots,L-1\}^D}
      q(r)\,
      \phi(y+a_k r).
    \label{eq:ordinary-block-spin-map}
  \end{equation}
  Periodic representatives are used when \(y+a_kr\) crosses the boundary.
  The normalization makes a constant microscopic field transform by the
  declared field-scaling factor \(L^{\Delta_\phi}\).
  \item A localization map which writes the pushforward
  \((B_k)_\#\mu_k\) in a Banach RG chart
  \[
    (g_k,K_k)\in E_{\rm loc}\oplus\mathcal K .
  \]
  The finite-dimensional coordinates \(g_k\) record relevant and marginal
  local terms.  The polymer activity \(K_k\) records the infinite-dimensional
  irrelevant tail in a norm controlling both spatial decay and large-field
  behavior.
\end{enumerate}
The structure is short-range only before the block step.  After integrating out
fluctuations, the effective interaction is generally not finite range; the
polymer norm and its contraction estimates are the mathematical replacement
for the microscopic finite-range property.
\end{definition}

\begin{construction}[Block-spin cumulants]
\label{constr:short-range-block-spin-reconstruction}
Let \(f\in C_c^\infty(\mathbb R^D)\).  The normalized lattice field at
spacing \(a_k\) is the random variable
\begin{equation}
  \Phi_k(f)
  =
  a_k^D\,a_k^{-\Delta_\phi}
  \sum_{x\in\Lambda_k} f(x)\phi_x .
  \label{eq:block-spin-distribution-pairing}
\end{equation}
For labels \(f_1,\ldots,f_n\), define the connected reconstructed
distribution by the joint cumulant
\begin{equation}
  \left\langle S_{n,k},f_1\otimes\cdots\otimes f_n\right\rangle
  =
  \kappa_{\mu_k}\bigl(
    \Phi_k(f_1),\ldots,\Phi_k(f_n)
  \bigr).
  \label{eq:block-spin-cumulant-distribution}
\end{equation}
The cumulant is computed from
\[
  \log
  \mathbb E_{\mu_k}
  \exp\!\left(\sum_{i=1}^n t_i\Phi_k(f_i)\right)
\]
by taking \(\partial_{t_1}\cdots\partial_{t_n}\) at the origin.  The
normalization in \eqref{eq:block-spin-distribution-pairing} is fixed by the
source pairing: if \(x=Ly\), then a source coupled to a field of dimension
\(\Delta_\phi\) scales with exponent \(D-\Delta_\phi\), as in
\eqref{eq:source-exponent-operator-dimension}.

The ordinary short-range scalar reconstruction problem is to prove, for
each \(n\), convergence of \(S_{n,k}\) in \(\mathcal S'((\mathbb R^D)^n)\)
or in a declared local distribution topology.  A Wilsonian RG proof supplies
this convergence through estimates in the RG chart:
\begin{equation}
  p_B\!\left(S_{n,k}-S_{n,\ast}\right)
  \leq
  C_{n,B}
  \bigl(
    \|g_k-g_\ast\|_{E_{\rm loc}}
    +\|K_k-K_\ast\|_{\mathcal K}
  \bigr)
  +\epsilon_{n,B,k},
  \label{eq:short-range-scalar-reconstruction-bound}
\end{equation}
where \(B\subset\mathcal S((\mathbb R^D)^n)\) is bounded and
\(\epsilon_{n,B,k}\to0\) is the discretization and finite-volume error.
If the RG trajectory is tuned so that
\begin{equation}
  \|g_k-g_\ast\|_{E_{\rm loc}}+\|K_k-K_\ast\|_{\mathcal K}
  \leq C_0\theta^k,\qquad 0<\theta<1 ,
  \label{eq:short-range-scalar-rg-chart-decay}
\end{equation}
then \eqref{eq:short-range-scalar-reconstruction-bound} is precisely a
reconstruction estimate of the form
\eqref{eq:rg-reconstruction-estimate}.  The point is that
\eqref{eq:short-range-scalar-rg-chart-decay} is the hard RG estimate, not
the vocabulary of fixed points or universality.
\end{construction}

\paragraph{Finite source windows to cumulant distributions.}
The cumulant notation in
\eqref{eq:block-spin-cumulant-distribution} is useful only after the
source-extension estimates have been connected to actual test-function
windows.  Fix a finite-dimensional test space
\[
  E=\operatorname{span}\{f_1,\ldots,f_m\}
  \subset C_c^\infty(\mathbb R^D)
\]
and write \(J_z=\sum_{a=1}^m z_a f_a\).  At scale \(k\), define the
finite-window connected source functional
\begin{equation}
  W_{k,E}(z)
  =
  \log
  \mathbb E_{\mu_k}\exp\{\Phi_k(J_z)\},
  \qquad z\in\mathbb C^m,
  \label{eq:short-range-source-window-functional}
\end{equation}
whenever the expectation is finite and holomorphic in the chosen polydisc.
If
\[
  \beta=(\beta_1,\ldots,\beta_m),
  \qquad
  |\beta|=\beta_1+\cdots+\beta_m,
\]
then
\begin{equation}
  \partial_z^\beta W_{k,E}(0)
  =
  \kappa_{\mu_k}\bigl(
    \underbrace{\Phi_k(f_1),\ldots,\Phi_k(f_1)}_{\beta_1\ {\rm times}},
    \ldots,
    \underbrace{\Phi_k(f_m),\ldots,\Phi_k(f_m)}_{\beta_m\ {\rm times}}
  \bigr).
  \label{eq:short-range-source-window-cumulant-derivative}
\end{equation}
This identity is just the definition of joint cumulants through the
logarithm of the moment-generating function, but it fixes the exact
derivative whose convergence is needed for the Schwinger distribution
window.

Suppose now that \(W_{\ast,E}\) is holomorphic on the same closed polydisc
\(P_\rho=\{|z_a|\le\rho_a\}\) and that
\begin{equation}
  \sup_{z\in P_\rho}|W_{k,E}(z)-W_{\ast,E}(z)|
  \leq \epsilon_{E,k},
  \qquad \epsilon_{E,k}\longrightarrow0 .
  \label{eq:short-range-source-window-uniform-convergence}
\end{equation}
The multidimensional Cauchy formula gives, for every multi-index \(\beta\),
\begin{equation}
  \left|
    \partial_z^\beta W_{k,E}(0)
    -
    \partial_z^\beta W_{\ast,E}(0)
  \right|
  \leq
  \frac{\beta!}{\rho^\beta}\,\epsilon_{E,k},
  \qquad
  \rho^\beta=\prod_a\rho_a^{\beta_a}.
  \label{eq:short-range-source-window-cauchy-cumulant-bound}
\end{equation}
Thus finite-source-window convergence in a complex neighborhood gives
quantitative convergence of every cumulant whose test functions lie in
\(E\).  If \(E\subset F\), compatibility of the source windows means
\[
  W_{k,F}(z_E,0)=W_{k,E}(z_E),
  \qquad
  W_{\ast,F}(z_E,0)=W_{\ast,E}(z_E),
\]
so the cumulants extracted from larger windows restrict to the smaller ones.

The passage from these finite windows to a distribution is a further
continuity statement, not a formal consequence of
\eqref{eq:short-range-source-window-cauchy-cumulant-bound}.  For each
\(n\), the RG construction must also provide an integer \(r_n\) and a
constant \(C_n\), independent of the finite test space \(E\) and of \(k\)
after the stated scale threshold.  Applied to any ordered \(n\)-tuple
\((h_1,\ldots,h_n)\subset E\), with source coordinates \(z_1,\ldots,z_n\)
chosen in those directions, the required bound is
\begin{equation}
  \left|
    \partial_{z_1}\cdots\partial_{z_n}
    W_{k,E}(0)
  \right|
  \leq
  C_n\,
  q_{r_n}(h_1\otimes\cdots\otimes h_n)
  \label{eq:short-range-source-window-uniform-schwartz-bound}
\end{equation}
for the declared Schwartz seminorm \(q_{r_n}\) on
\(\mathcal S((\mathbb R^D)^n)\), and the same bound for the limiting window.
Compatibility plus
\eqref{eq:short-range-source-window-uniform-schwartz-bound} is precisely the
input of Lemma~\ref{lem:rg-projective-window-extension-distribution}.  It
produces the limiting tempered distribution \(S_{n,\ast}\); the Cauchy
estimate
\eqref{eq:short-range-source-window-cauchy-cumulant-bound} supplies the
finite-window convergence rate.  Local source counterterms are handled as
additional polynomial source coordinates whose derivatives are distributions
supported on collision diagonals.  They are part of the source window, not an
after-the-fact modification of separated cumulants.

\paragraph{Source majorants for the OS-II growth input.}
The preceding paragraph controls fixed cumulant windows.  The corrected
OS-II reconstruction theorem needs a uniform many-insertion growth estimate
for the Schwinger hierarchy itself, not only convergence at each fixed
\(n\).  The corresponding Wilsonian estimate is most naturally stated for
the moment source functional
\begin{equation}
  Z_{k,E}(z)
  =
  \mathbb E_{\mu_k}\exp\{\Phi_k(J_z)\},
  \qquad
  W_{k,E}(z)=\log Z_{k,E}(z),
  \label{eq:short-range-moment-source-window}
\end{equation}
on the same finite test space \(E\).  Let
\[
  N_n=N_0+C_{\rm OS}n
\]
be the seminorm order allowed in the OS-II linear-growth condition, and let
\(q_{1,N_n}\) be the corresponding one-variable Schwartz seminorm.  A
source-window estimate strong enough for OS-II has the following form.  For
every ordered \(n\)-tuple \(h_1,\ldots,h_n\) in the test space, the functional
\(Z_{k,E}\), \(E=\operatorname{span}\{h_1,\ldots,h_n\}\), is holomorphic on
the polydisc
\[
  |z_i|\leq \frac{\rho}{q_{1,N_n}(h_i)},\qquad i=1,\ldots,n,
\]
with constants \(\rho,A,B,\gamma>0\), independent of \(k,n\), and of the
chosen tuple, such that
\begin{equation}
  \sup_{P_\rho(h_1,\ldots,h_n)}
  |Z_{k,E}(z)|
  \leq
  A\,B^n\,(n!)^\gamma .
  \label{eq:short-range-source-majorant-osii}
\end{equation}
Cauchy's formula in the \(n\) independent source variables then gives
\begin{equation}
  |S_{n,k}(h_1\otimes\cdots\otimes h_n)|
  \leq
  A\left(\frac B\rho\right)^n(n!)^\gamma
  \prod_{i=1}^n q_{1,N_n}(h_i).
  \label{eq:short-range-source-majorant-cauchy-moment}
\end{equation}
For a general zero-diagonal test function \(F\), let \(\pi_{n,N_n}(F)\) be
the projective tensor seminorm obtained by decomposing \(F\) into finite sums
of simple tensors and taking the infimum of the corresponding products of
one-variable \(q_{1,N_n}\)-seminorms.  Linearity gives
\begin{equation}
  |S_{n,k}(F)|
  \leq
  A\left(\frac B\rho\right)^n(n!)^\gamma
  \pi_{n,N_n}(F).
  \label{eq:short-range-source-majorant-projective-bound}
\end{equation}
Thus the OS-II bound
\eqref{eq:os-linear-growth-condition} follows from the source majorant
whenever the chosen Schwartz seminorm family satisfies a comparison estimate
\[
  \pi_{n,N_n}(F)
  \leq
  C_\pi^n q_{n,N'_0+C'_{\rm OS}n}(F)
\]
on the zero-diagonal test class.  The constants in the OS-II growth condition
are then \(B_{\rm OS}=C_\pi B/\rho\), with the same factorial exponent
\(\gamma\).  This is a stricter demand than fixed-\(n\) source convergence:
if the available source radius shrinks with \(n\), if the supremum in
\eqref{eq:short-range-source-majorant-osii} has uncontrolled combinatorial
growth, or if only connected cumulants are bounded without a summable
partition estimate, the OS-II analytic input has not been obtained.

\paragraph{From connected cumulants to moments.}
The source functional most directly produced by a polymer or cluster RG
construction is often \(W_{k,E}=\log Z_{k,E}\), hence connected cumulants.
The corrected OS-II growth hypothesis is a bound on Schwinger functions
themselves.  The passage from one to the other is the moment--cumulant
partition sum, and its constants must be controlled uniformly in the number
of insertions.

\begin{lemma}[Cumulant partition bound for OS-II moments]
\label{lem:rg-cumulant-partition-osii-bound}
Fix \(n\geq1\), test functions \(h_1,\ldots,h_n\), and the seminorm
\(q_{1,N_n}\) used in the preceding paragraph.  For each nonempty
\(B\subset\{1,\ldots,n\}\), let
\(\kappa_B(h_B)\) be the connected cumulant with insertion set \(B\), where
\(h_B=(h_i)_{i\in B}\).  Assume that constants
\(A_{\rm c},C_{\rm c},\gamma_{\rm c}\), independent of \(n\) and of the
chosen test functions, satisfy
\begin{equation}
  |\kappa_B(h_B)|
  \leq
  A_{\rm c}\,C_{\rm c}^{|B|}\,(|B|!)^{\gamma_{\rm c}}
  \prod_{i\in B}q_{1,N_n}(h_i)
  \label{eq:rg-connected-cumulant-osii-input}
\end{equation}
for every nonempty \(B\).  Let
\[
  M_n(h_1,\ldots,h_n)
  =
  \sum_{\pi\in\Pi_n}\prod_{B\in\pi}\kappa_B(h_B)
\]
be the corresponding Schwinger moment, with \(\Pi_n\) the set of set
partitions of \(\{1,\ldots,n\}\).  Then
\begin{equation}
  |M_n(h_1,\ldots,h_n)|
  \leq
  \bigl(2C_{\rm c}\max(1,A_{\rm c})\bigr)^n
  (n!)^{\gamma_{\rm c}+1}
  \prod_{i=1}^n q_{1,N_n}(h_i).
  \label{eq:rg-connected-to-moment-osii-bound}
\end{equation}
\end{lemma}

\begin{proof}
The moment--cumulant identity is the finite identity obtained by expanding
\(\exp W\).  Taking absolute values and applying
\eqref{eq:rg-connected-cumulant-osii-input} gives
\[
  |M_n(h_1,\ldots,h_n)|
  \leq
  C_{\rm c}^{n}
  \prod_{i=1}^n q_{1,N_n}(h_i)
  \sum_{\pi\in\Pi_n}
  A_{\rm c}^{|\pi|}
  \prod_{B\in\pi}(|B|!)^{\gamma_{\rm c}} .
\]
For every partition \(\pi\), one has
\(\prod_{B\in\pi}|B|!\leq n!\), since the left hand side counts
permutations internal to the blocks and is bounded by all permutations of
the \(n\) labels.  Also
\(A_{\rm c}^{|\pi|}\leq \max(1,A_{\rm c})^n\).  Finally,
\(|\Pi_n|\leq 2^{n-1}n!\): every partition can be obtained by choosing a
permutation of the \(n\) labels and then choosing cuts between consecutive
labels, so the stated number is a surjective overcount.  Combining these
three bounds gives \eqref{eq:rg-connected-to-moment-osii-bound}.
\end{proof}

Consequently a connected-cumulant RG theorem can feed the corrected OS-II
hypothesis only after the constants in
\eqref{eq:rg-connected-cumulant-osii-input} are obtained uniformly in the
insertion number and with the declared seminorm order \(N_n=N_0+C_{\rm OS}n\).
The estimate loses one factorial power when no sharper partition control is
available.  This loss is part of the theorem statement, not a matter of
notation: if the connected bound carries constants depending on the number
of clusters, on the source-window dimension, or on an \(n\)-dependent radius,
then \eqref{eq:rg-connected-to-moment-osii-bound} does not supply the OS-II
growth input.

\paragraph{RG-chart control of a holomorphic source window.}
The hypothesis
\eqref{eq:short-range-source-window-uniform-convergence} is not supplied by
knowing that the unsourced local coordinates converge.  A Wilsonian proof
must control the source-extended chart itself.  For a fixed finite test
space \(E\) and a fixed closed polydisc \(P_\rho\), the source extension of
the RG chart should give a decomposition of the connected source functional
of the form
\begin{equation}
  W_{k,E}(z)
  =
  C_{k,E}(z)
  +
  \mathfrak W_E\bigl(g_{k,E}(z),K_{k,E}(z)\bigr)
  +
  r_{k,E}(z),
  \qquad z\in P_\rho .
  \label{eq:short-range-source-window-rg-chart-decomposition}
\end{equation}
Here \(C_{k,E}\) is the local normalizing coordinate, including source
dependent vacuum and contact coordinates, \(g_{k,E}\) are the retained local
coordinates, \(K_{k,E}\) is the source-decorated polymer tail, and
\(r_{k,E}\) is the declared sampling, boundary, and finite-step remainder.
The map \(\mathfrak W_E\) is the finite-window evaluation map of the RG
chart.  It is not assumed to be a universal function independent of the
chosen chart; it is part of the reconstruction datum for this finite source
window.

Suppose that all chart points occurring in
\eqref{eq:short-range-source-window-rg-chart-decomposition} lie in a common
tube \(\mathcal T_E\) on which \(\mathfrak W_E\) obeys
\begin{equation}
  \left|
    \mathfrak W_E(g,K)-\mathfrak W_E(g',K')
  \right|
  \leq
  L_E\bigl(\|g-g'\|_{E_{\rm loc}}+\|K-K'\|_{\mathcal K_E}\bigr).
  \label{eq:short-range-source-window-chart-lipschitz}
\end{equation}
If limiting source-chart data
\((C_{\ast,E},g_{\ast,E},K_{\ast,E},r_{\ast,E})\) are holomorphic on the same
polydisc and
\begin{align}
  \delta^C_{E,k}
  &:=
  \sup_{P_\rho}|C_{k,E}-C_{\ast,E}|,\\
  \delta^g_{E,k}
  &:=
  \sup_{P_\rho}\|g_{k,E}-g_{\ast,E}\|_{E_{\rm loc}},\\
  \delta^K_{E,k}
  &:=
  \sup_{P_\rho}\|K_{k,E}-K_{\ast,E}\|_{\mathcal K_E},\\
  \delta^r_{E,k}
  &:=
  \sup_{P_\rho}|r_{k,E}-r_{\ast,E}|
\end{align}
all tend to zero, then
\begin{equation}
  \sup_{z\in P_\rho}|W_{k,E}(z)-W_{\ast,E}(z)|
  \leq
  \delta^C_{E,k}
  +
  L_E(\delta^g_{E,k}+\delta^K_{E,k})
  +
  \delta^r_{E,k}.
  \label{eq:short-range-source-window-chart-to-holomorphic-bound}
\end{equation}
Inserting this bound into
\eqref{eq:short-range-source-window-cauchy-cumulant-bound} gives the
cumulant convergence rate for the finite window.  The fixed polydisc is
load-bearing: a bound on \(P_{\rho_k}\) with \(\rho_k\downarrow0\) does not
give a uniform Cauchy estimate for derivatives.  Thus source-radius control,
local normalizing coordinates, source-decorated irrelevant tails, and
finite-step remainders are all part of the Wilsonian reconstruction theorem;
they cannot be recovered from an unsourced fixed-point statement.

\paragraph{Source-stable trajectory estimate.}
The source-extended chart also has its own stable and unstable directions.
This is where local contact terms enter the RG theorem rather than being
added after the continuum limit.  For a finite test space \(E\), write
\[
  g_{k,E}(z)-g_{\ast,E}(z)
  =
  u_{k,E}(z)+s_{k,E}(z)
\]
according to the unstable/retained and stable/irrelevant splitting of the
source-local coordinate space.  The unstable part includes ordinary
relevant couplings and any source-local or contact coordinates whose
engineering or anomalous exponent makes them non-decaying in the source
window.  For a fixed polydisc \(P_\rho\), a finite-depth RG estimate has the
form
\begin{align}
  \sup_{P_\rho}\|s_{k+1,E}\|
  &\le
  \theta_E\sup_{P_\rho}\|s_{k,E}\|+\varepsilon^s_{k,E},
  \qquad 0<\theta_E<1,
  \label{eq:source-stable-chart-recursion}\\
  \sup_{P_\rho}\|u_{k+1,E}\|
  &\le
  \Lambda_E\sup_{P_\rho}\|u_{k,E}\|+\varepsilon^u_{k,E},
  \qquad \Lambda_E>1 .
  \label{eq:source-unstable-chart-recursion}
\end{align}
The first recursion gives
\begin{equation}
  \sup_{P_\rho}\|s_{N,E}\|
  \le
  \theta_E^N\sup_{P_\rho}\|s_{0,E}\|
  +
  \sum_{j=0}^{N-1}\theta_E^{\,N-1-j}\varepsilon^s_{j,E}.
  \label{eq:source-stable-chart-telescoping}
\end{equation}
The second recursion gives the finite-depth amplification bound
\begin{equation}
  \sup_{P_\rho}\|u_{N,E}\|
  \le
  \Lambda_E^N\sup_{P_\rho}\|u_{0,E}\|
  +
  \sum_{j=0}^{N-1}\Lambda_E^{\,N-1-j}\varepsilon^u_{j,E}.
  \label{eq:source-unstable-chart-amplification}
\end{equation}
Thus a source-local coordinate whose unstable component is not tuned to
order \(\Lambda_E^{-N}\), up to the accumulated source-local defects, can
spoil the finite-window limit even if the unsourced scalar trajectory is on
its critical surface.

Combining
\eqref{eq:source-stable-chart-telescoping}--\eqref{eq:source-unstable-chart-amplification}
with
\eqref{eq:short-range-source-window-chart-to-holomorphic-bound} gives the
actual holomorphic source-window error bound
\begin{equation}
  \epsilon_{E,N}^{\rm src}
  \le
  \delta^C_{E,N}
  +
  L_E\!\left(
    \sup_{P_\rho}\|u_{N,E}\|
    +
    \sup_{P_\rho}\|s_{N,E}\|
    +
    \delta^K_{E,N}
  \right)
  +
  \delta^r_{E,N}.
  \label{eq:source-stable-chart-to-window-bound}
\end{equation}
This is the source-window version of stable-manifold tuning.  The
source-local counterterms and collision-diagonal coordinates must either be
retained and tuned in \(u_{k,E}\), shown to contract in \(s_{k,E}\), or
bounded in the explicit remainder.  If they are simply omitted from the
chart, the Wilsonian RG estimate has not yet produced the Schwinger
distribution window used in OS reconstruction.

\paragraph{Finite volume to infinite volume for a source window.}
The thermodynamic step at fixed lattice spacing is another estimate, not a
notation change.  At fixed lattice spacing \(a_k\), let
\(\Omega\subset a_k\mathbb Z^D\) be a finite connected region and write
\(W_{k,\Omega,E}(z)\) for the connected source functional
\eqref{eq:short-range-source-window-functional} computed with the
finite-volume Gibbs measure on \(\Omega\) and the declared local boundary
convention.  Let
\[
  K_{E,k}
  =
  \Omega\cap
  \bigcup_{a=1}^m \operatorname{supp} f_a
\]
be the lattice source support, after the continuum test functions are
restricted to \(a_k\mathbb Z^D\), and let
\[
  R_\Omega=\operatorname{dist}_k(K_{E,k},a_k\mathbb Z^D\setminus\Omega)
\]
be the lattice distance from the source support to the boundary of
\(\Omega\), in physical units or in lattice units according to the declared
cluster norm.  Suppose the source-window cluster expansion gives, on a fixed
closed polydisc \(P_\rho\), the boundary-tail estimate
\begin{equation}
  \sup_{z\in P_\rho}
  |W_{k,\Omega',E}(z)-W_{k,\Omega,E}(z)|
  \leq
  A_E\,\Xi(R_\Omega)
  \label{eq:short-range-source-window-volume-tail}
\end{equation}
whenever \(\Omega\subset\Omega'\) and both regions contain \(K_{E,k}\), where
\(\Xi(R)\to0\) as \(R\to\infty\).  In a finite-range or exponentially
clustering polymer estimate one expects
\(\Xi(R)\le C\exp(-\zeta R)\); in a critical construction the available
\(\Xi\) is whatever the RG polymer norm actually proves.  Equation
\eqref{eq:short-range-source-window-volume-tail} is the statement that
clusters touching the source window and clusters that feel the boundary have
a summable connected bridge.

For every multi-index \(\beta\), Cauchy's formula on the same polydisc gives
\begin{equation}
  \left|
    \partial_z^\beta W_{k,\Omega',E}(0)
    -
    \partial_z^\beta W_{k,\Omega,E}(0)
  \right|
  \leq
  \frac{\beta!}{\rho^\beta}\,A_E\,\Xi(R_\Omega).
  \label{eq:short-range-source-window-volume-cauchy-bound}
\end{equation}
Thus the finite-volume cumulant windows are Cauchy as the boundary recedes
from the source support.  Their limit is denoted
\[
  \partial_z^\beta W_{k,\infty,E}(0)
  :=
  \lim_{\Omega\uparrow a_k\mathbb Z^D}
  \partial_z^\beta W_{k,\Omega,E}(0),
  \]
provided the directed family of regions is chosen so that
\(R_\Omega\to\infty\).  The resulting window is independent of the chosen
cofinal exhaustion whenever the same tail estimate controls any two
exhaustions through a common larger region.  This is the finite-volume part
of item 1 in
Definition~\ref{def:short-range-scalar-os-assembly-datum}.  Without a bound
of the form
\eqref{eq:short-range-source-window-volume-tail}, a sequence of finite
volumes may have plausible local numerical stability while failing to define
the thermodynamic source window required by OS reconstruction.

The independence of the exhaustion is also part of the same estimate.  Let
\(\{\Omega_i\}\) and \(\{\widetilde\Omega_j\}\) be two directed exhaustions
whose boundary distances from \(K_{E,k}\) tend to infinity.  For fixed
\(i,j\), choose a connected region \(\Omega_{ij}^\sharp\) containing both
\(\Omega_i\) and \(\widetilde\Omega_j\).  Applying
\eqref{eq:short-range-source-window-volume-tail} twice gives
\begin{equation}
  \sup_{z\in P_\rho}
  |W_{k,\Omega_i,E}(z)-W_{k,\widetilde\Omega_j,E}(z)|
  \leq
  A_E\Xi(R_{\Omega_i})+
  A_E\Xi(R_{\widetilde\Omega_j}).
  \label{eq:short-range-source-window-cofinal-volume-bound}
\end{equation}
The right side tends to zero as \(i,j\to\infty\), so the holomorphic limit on
\(P_\rho\) is independent of the cofinal volume sequence.  Cauchy's formula
then yields the derivative comparison
\begin{equation}
  \left|
    \partial_z^\beta W_{k,\Omega_i,E}(0)
    -
    \partial_z^\beta W_{k,\widetilde\Omega_j,E}(0)
  \right|
  \leq
  \frac{\beta!}{\rho^\beta}
  \bigl(
    A_E\Xi(R_{\Omega_i})+
    A_E\Xi(R_{\widetilde\Omega_j})
  \bigr).
  \label{eq:short-range-source-window-cofinal-derivative-bound}
\end{equation}

The same bookkeeping fixes the joint thermodynamic and continuum limit used
later in the RG-to-OS assembly package.  Suppose that, for the same finite
source window \(E\), the RG-chart comparison to a limiting source functional
\(W_{\ast,E}\) gives an error \(\epsilon^{\rm RG}_{E,k}\) on \(P_\rho\),
while the chosen finite volume \(\Omega_k\) obeys the boundary distance
\(R_{\Omega_k}\to\infty\).  Combining
\eqref{eq:short-range-source-window-chart-to-holomorphic-bound} with
\eqref{eq:short-range-source-window-volume-tail} gives
\begin{equation}
  \sup_{z\in P_\rho}
  |W_{k,\Omega_k,E}(z)-W_{\ast,E}(z)|
  \leq
  \epsilon^{\rm RG}_{E,k}
  +
  A_E\Xi(R_{\Omega_k}).
  \label{eq:short-range-source-window-joint-volume-rg-bound}
\end{equation}
Consequently every extracted cumulant in the \(E\)-window satisfies
\begin{equation}
  \left|
    \partial_z^\beta W_{k,\Omega_k,E}(0)
    -
    \partial_z^\beta W_{\ast,E}(0)
  \right|
  \leq
  \frac{\beta!}{\rho^\beta}
  \left(
    \epsilon^{\rm RG}_{E,k}
    +
    A_E\Xi(R_{\Omega_k})
  \right).
  \label{eq:short-range-source-window-joint-volume-rg-derivative-bound}
\end{equation}
Thus the finite-volume simulation or finite-volume constructive estimate may
be used as thermodynamic Schwinger data only through a boundary-tail function
and a scale schedule that makes the combined right side tend to zero for
each fixed source window.

\paragraph{Adjoint source blocking and smooth-test error.}
The source normalization in
\eqref{eq:block-spin-distribution-pairing} is not a convention that can be
chosen independently at each scale.  Let \(F\) be a real-valued test
function on the coarse lattice \(\Lambda_{k-1}\).  The source paired with
the blocked field is
\[
  a_{k-1}^{D-\Delta_\phi}\sum_{y\in\Lambda_{k-1}}
  F(y)(B_k\phi)(y).
\]
Using \(a_{k-1}=La_k\) and the definition
\eqref{eq:ordinary-block-spin-map}, this expression is equal to
\begin{equation}
  a_k^{D-\Delta_\phi}
  \sum_{y\in\Lambda_{k-1}}
  \sum_{r\in\{0,\ldots,L-1\}^D}
  L^D q(r)F(y)\,\phi_{y+a_kr}.
  \label{eq:adjoint-source-blocking-derivation}
\end{equation}
Thus the adjoint source map associated with the normalized pairings is
\begin{equation}
  (B_k^\dagger F)(y+a_kr)=L^D q(r)F(y),
  \qquad r\in\{0,\ldots,L-1\}^D,
  \label{eq:adjoint-source-blocking-map}
\end{equation}
with periodic representatives at the boundary.  Equivalently,
\begin{equation}
  a_{k-1}^{D-\Delta_\phi}\sum_y F(y)(B_k\phi)(y)
  =
  a_k^{D-\Delta_\phi}\sum_x (B_k^\dagger F)(x)\phi_x .
  \label{eq:adjoint-source-blocking-identity}
\end{equation}
The exponent \(D\), rather than \(D-\Delta_\phi\), in
\eqref{eq:adjoint-source-blocking-map} is forced by the factor
\(L^{\Delta_\phi}\) already present in the blocked field.  This is the
finite-regulator identity behind source-window extraction in an ordinary
short-range scalar RG chart.

For comparison with continuum test functions, take the uniform kernel
\(q(r)=L^{-D}\) and let \(f\in C_c^1(\mathbb R^D)\).  Put \(F(y)=f(y)\) on
the coarse lattice and let \(Y_k\) be a finite set of coarse sites containing
all blocks that meet the support of \(f\).  Then
\begin{equation}
  a_k^D
  \sum_{y\in Y_k}
  \sum_{r\in\{0,\ldots,L-1\}^D}
  \left|(B_k^\dagger F)(y+a_kr)-f(y+a_kr)\right|
  \le
  C_{D,L}\,\|\nabla f\|_\infty\,a_{k-1}\,
  \operatorname{Vol}_{k-1}(Y_k),
  \label{eq:adjoint-source-smooth-test-error}
\end{equation}
where
\[
  \operatorname{Vol}_{k-1}(Y_k)=a_{k-1}^D|Y_k|,
  \qquad
  C_{D,L}=L^{-(D+1)}
  \sum_{r\in\{0,\ldots,L-1\}^D}|r|.
\]
Indeed, for the uniform kernel \(B_k^\dagger F\) is constant on each block
and equal to \(f(y)\), while
\(|f(y)-f(y+a_kr)|\leq a_k|r|\,\|\nabla f\|_\infty\).  Multiplication by
\(a_k^D\) and summation over \(y\) and \(r\) gives
\eqref{eq:adjoint-source-smooth-test-error}.  Higher-order block kernels may
improve this sampling error when their moments are matched to the chosen
cell coordinates; without those moment conditions the exact identity
\eqref{eq:adjoint-source-blocking-identity}, not pointwise approximation of
smooth tests, is the invariant source statement.

This construction also fixes the role of the block kernel.  A different
kernel \(q\), a different field-strength normalization, or a different
localization map may give an equivalent continuum theory only after a
comparison map with estimates in the sense of
Definition~\ref{def:comparison-of-rg-descriptions} is supplied.  Finite-range
microscopic interactions alone do not identify the continuum normalization
of the field.

\begin{lemma}[Reflection-positive block-spin pullback]
\label{lem:reflection-positive-block-spin-pullback}
Let \(\Lambda_k\) be split by a time reflection \(\theta\) into positive,
negative, and fixed-time sites, and let \(\mathcal A_{k,+}\) be the algebra
of polynomials in the positive-time fine fields.  Suppose the finite measure
\(\mu_k\) is reflection positive:
\[
  \int_{\mathbb R^{\Lambda_k}}
  \overline{\theta P(\phi)}\,P(\phi)\,\dd\mu_k(\phi)\ge0,
  \qquad P\in\mathcal A_{k,+}.
\]
Assume that the block-spin map \(B_k\) is reflection compatible in the
following finite sense: every coarse positive-time field
\(\Phi_y=(B_k\phi)(y)\), \(y^0>0\), is a real linear combination of fine
fields in the positive-time half-lattice, and
\((B_k\phi)(\theta y)=(B_k(\theta\phi))(y)\).  Let
\(\nu_{k-1}:=(B_k)_\#\mu_k\) be the blocked measure.  Then \(\nu_{k-1}\) is
reflection positive for the coarse positive-time polynomial algebra.
\end{lemma}

\begin{proof}
Let \(F\) be a polynomial in coarse positive-time fields, and define its
pullback
\[
  B_k^\ast F(\phi)=F(B_k\phi).
\]
By the support assumption on the positive-time block stencils,
\(B_k^\ast F\in\mathcal A_{k,+}\).  Reflection compatibility gives
\[
  \theta(B_k^\ast F)=B_k^\ast(\theta F),
\]
because reflecting the fine configuration and then blocking is the same, on
the reflected coarse lattice, as blocking and then reflecting.  Therefore
\[
\begin{aligned}
  \int \overline{\theta F(\Phi)}\,F(\Phi)\,\dd\nu_{k-1}(\Phi)
  &=
  \int \overline{B_k^\ast(\theta F)(\phi)}\,
        B_k^\ast F(\phi)\,\dd\mu_k(\phi)  \\
  &=
  \int \overline{\theta(B_k^\ast F)(\phi)}\,
        B_k^\ast F(\phi)\,\dd\mu_k(\phi)
  \ge0 .
\end{aligned}
\]
Thus every finite coarse reflection-positivity Gram matrix is the compression
of a fine reflection-positivity Gram form under the pullback \(B_k^\ast\).
The lemma is finite-regulator only.  A continuum OS statement still requires
the directed-family convergence and growth estimates described in
Definition~\ref{def:local-qft-strength-observable-germ}.
\end{proof}

\paragraph{Requirements for OS reconstruction.}
The preceding construction and lemma are finite pieces.  A critical
short-range scalar limit reaches a local QFT through a directed family of
estimates
in exactly the form required by
Definition~\ref{def:os-admissible-euclidean-data}.

\begin{definition}[Short-range scalar OS reconstruction requirements]
\label{def:short-range-scalar-os-assembly-datum}
Fix a short-range scalar block-spin structure as in
Definition~\ref{def:short-range-scalar-block-spin-structure} and a tuned
trajectory \(V_k=(g_k,K_k)\) in its RG chart.  A short-range scalar
OS reconstruction argument consists of the following additional objects and
estimates.
\begin{enumerate}
  \item An infinite-volume prescription for the cumulants
  \(S_{n,k}\).  Either the finite-volume limit is taken at fixed lattice
  spacing before \(k\to\infty\), or a joint volume--continuum limit is
  specified, with estimates uniform enough to give the same limiting
  distributions on every compactly supported test window.
  \item Distributional reconstruction estimates: for every \(n\) and every
  bounded set \(B\subset\mathcal S((\mathbb R^D)^n)\),
  \[
    p_B(S_{n,k}-S_{n,\ast})\longrightarrow0 ,
  \]
  together with a uniform temperedness bound of the form
  \[
    |S_{n,k}(f)|
    \le C_n\,q_{n,r_n}(f),
    \qquad f\in\mathcal S((\mathbb R^D)^n),
  \]
  for a declared Schwartz seminorm \(q_{n,r_n}\).  This bound is the
  continuity estimate that turns a projective family of finite windows into
  a tempered distribution, as in
  Lemma~\ref{lem:rg-projective-window-extension-distribution}.
  \item Euclidean covariance, permutation symmetry, normalization
  \(S_{0,k}=1\), and Hermiticity at finite regulator, or regulator defects
  for these identities that vanish in the same distributional topology.
  \item Reflection positivity in a directed positive-time test family.  More
  precisely, for every finite positive-time polynomial test family
  \(\mathcal F=\{F_1,\ldots,F_m\}\), the regulated Gram matrices
  \[
    G_{k,\mathcal F;ij}
    =
    S_k((\Theta F_i)F_j)
  \]
  are positive semidefinite, or obey a defect inequality whose defect tends
  to zero, and their entries converge to the limiting Gram matrix.  If one
  uses quantitative entrywise error estimates for increasing families, the
  estimates must be compatible with the family size in the sense explained
  below \eqref{eq:rg-finite-os-positivity-bound}.
  \item Positive-time semigroup regularity on the OS quotient.  This means
  that the positive-time translations defined from the limiting Schwinger
  hierarchy obey the semigroup law and strong continuity conditions
  \eqref{eq:os-time-semigroup-law}--\eqref{eq:os-time-strong-continuity}.
  In an RG proof this is an estimate on translated positive-time test
  windows, additional to the finite Gram-form positivity estimates.
  \item Corrected OS-II analytic regularity.  A sufficient form is the
  uniform linear-growth estimate
  \[
    |S_{n,k}^{\rm red}(F)|
    \le
    A\,B^n\,(n!)^\gamma\,
    q_{n,N_0+C_{\rm OS}n}(F)
  \]
  for all sufficiently large \(k\), with constants
  \(A,B,\gamma,N_0,C_{\rm OS}\) independent of \(n\) and \(k\).  The same
  bound then passes to the limiting reduced hierarchy by distributional
  convergence.  Equivalently, one may provide a
  different model-specific analytic theorem that yields the Lorentzian
  boundary-value package of
  Definition~\ref{def:os-lorentzian-boundary-value-package}.  Ordinary
  temperedness at each fixed \(n\) supplies fixed-\(n\) continuity; the
  OS-II item controls the many-variable analytic continuation uniformly in
  the insertion number.  In a Wilsonian source construction, a usable route
  to this item is a moment-source majorant of the form
  \eqref{eq:short-range-source-majorant-osii}, together with the
  projective-seminorm comparison following
  \eqref{eq:short-range-source-majorant-projective-bound}.
  \item The ordered-insertion adjoint identity required in
  Theorem~\ref{thm:os-reconstruction}; for a single real scalar field this is
  the compatibility of Hermiticity with the positive-time multiplication
  convention, while for a labelled field family it is an identity of the
  declared label reflection.
\end{enumerate}
Clustering estimates are added when the target claim includes
uniqueness of the vacuum.
\end{definition}

\paragraph{Positive-time translation continuity from finite windows.}
Item 5 of
Definition~\ref{def:short-range-scalar-os-assembly-datum} is a continuity
estimate in the limiting OS Hilbert norm.  That norm is not the Schwartz norm
of a Euclidean test function; it is the limiting positive-time Gram norm.
Thus strong continuity of the positive-time translation semigroup is proved
by estimates on translated Gram windows.

Let \(F\) be a polynomial positive-time test vector whose Euclidean-time
support has distance \(d_F>0\) from the reflection plane.  For
\(0\leq t<d_F/2\), let \(\tau_tF\) denote translation of every Euclidean-time
argument of \(F\) forward by \(t\).  Suppose the regulated Schwinger
hierarchy gives, uniformly for all sufficiently large \(k\),
\begin{equation}
  S_k\!\left(
    \Theta(\tau_tF-F)(\tau_tF-F)
  \right)
  \leq \omega_F(t)+\varepsilon_{F,k},
  \qquad
  \lim_{t\downarrow0}\omega_F(t)=0,\qquad
  \varepsilon_{F,k}\to0 .
  \label{eq:rg-positive-time-translation-window-bound}
\end{equation}
Here the left side is a finite positive-time Gram-window entry, and the
support restriction ensures that \(\tau_tF-F\) still belongs to the
positive-time test algebra.  If the corresponding Gram entries converge to
the limiting OS form, then
\[
  \|[\tau_tF]-[F]\|_{\mathcal H_{\rm OS}}^2
  =
  S_\ast\!\left(
    \Theta(\tau_tF-F)(\tau_tF-F)
  \right)
  \leq \omega_F(t),
\]
after taking \(k\to\infty\).  Therefore \(t\mapsto[\tau_tF]\) is continuous at
\(t=0\) on every such vector.  Translation covariance gives
\(\tau_t\tau_sF=\tau_{t+s}F\) while the support remains positive-time; after
passing to the OS quotient this supplies the semigroup law on the dense
positive-time domain.  Extending the semigroup from this domain to the OS
Hilbert-space closure requires the same bound on a directed family of
positive-time vectors whose classes are dense.  A Wilsonian RG proof must
therefore supply \eqref{eq:rg-positive-time-translation-window-bound}, or an
equivalent model-specific equicontinuity estimate, in addition to finite
reflection positivity.

\begin{lemma}[Positive-time quotient semigroup criterion]
\label{lem:rg-positive-time-quotient-semigroup-criterion}
Let \(P_\infty\) be the algebraic positive-time test space and let
\[
  Q(F,G)=S_\ast((\Theta F)G)
\]
be the limiting positive semidefinite OS form on \(P_\infty\).  Put
\[
  \mathcal N_Q=\{F\in P_\infty\mid Q(F,F)=0\},
  \qquad
  \|[F]\|_{\rm OS}^2=Q(F,F).
\]
Assume that for \(0\le t\le t_0\) the positive-time translations
\(\tau_t:P_\infty\to P_\infty\) satisfy \(\tau_0=1\) and
\(\tau_t\tau_s=\tau_{t+s}\) whenever \(s,t,s+t\in[0,t_0]\).  Assume also the
quotient-stability estimate
\begin{equation}
  Q(\tau_tF,\tau_tF)\le C_t\,Q(F,F),
  \qquad C_t<\infty ,
  \label{eq:rg-os-quotient-null-stability-bound}
\end{equation}
with \(\sup_{0\le t\le t_0}C_t<\infty\), and the limiting Gram-continuity
estimate
\[
  Q(\tau_tF-F,\tau_tF-F)\longrightarrow0
  \qquad (t\downarrow0)
\]
for every \(F\in P_\infty\).  Then
\[
  T_t[F]=[\tau_tF]
\]
is a well-defined strongly continuous local semigroup on the OS Hilbert-space
completion of \(P_\infty/\mathcal N_Q\) for \(0\le t\le t_0\), with
\(\|T_t\|\le C_t^{1/2}\) on its initial domain of definition.  If the same
estimates hold on every compact positive-time interval, these local semigroups
assemble to a strongly continuous semigroup for \(t\ge0\).  If \(C_t\le1\),
the corresponding semigroup is contractive.
\end{lemma}

\begin{proof}
The first load-bearing point is descent to the quotient.  If
\(F\in\mathcal N_Q\), then \eqref{eq:rg-os-quotient-null-stability-bound} gives
\[
  Q(\tau_tF,\tau_tF)\le C_tQ(F,F)=0,
\]
so \(\tau_tF\in\mathcal N_Q\).  Hence the class \([\tau_tF]\) depends only on
the class \([F]\).  The same estimate gives
\[
  \|T_t[F]\|_{\rm OS}^2=Q(\tau_tF,\tau_tF)
  \le C_tQ(F,F)=C_t\|[F]\|_{\rm OS}^2,
\]
so \(T_t\) is bounded on the quotient norm and extends uniquely to the Hilbert
completion.

The semigroup law is inherited from the algebraic identity
\(\tau_t\tau_sF=\tau_{t+s}F\) on the common positive-time domain and then
extends by continuity.  It remains to prove strong continuity on the
completion.  On the dense quotient \(P_\infty/\mathcal N_Q\), the assumed
Gram-continuity estimate is exactly
\[
  \|T_t[F]-[F]\|_{\rm OS}^2
  =
  Q(\tau_tF-F,\tau_tF-F)\longrightarrow0 .
\]
For a general Hilbert vector \(v\), choose \([F]\) with
\(\|v-[F]\|_{\rm OS}<\alpha\).  If \(M=\sup_{0\le t\le t_0}C_t^{1/2}\), then
\[
  \|T_tv-v\|_{\rm OS}
  \le (M+1)\alpha+\|T_t[F]-[F]\|_{\rm OS}.
\]
First choose \(\alpha\) small and then \(t\) small.  This proves strong
continuity at \(t=0\), and the semigroup law gives strong continuity at every
positive time in the interval.
\end{proof}

This lemma isolates an estimate that is logically prior to writing
\([\tau_tF]\) as an OS Hilbert-space vector.  The finite-window bound
\eqref{eq:rg-positive-time-translation-window-bound} supplies the continuity
term after the regulator is removed.  The quotient-stability bound
\eqref{eq:rg-os-quotient-null-stability-bound} supplies invariance of the OS
null space.  In a model proof both estimates should be derived from the same
directed Gram-window control.

Given Definition~\ref{def:short-range-scalar-os-assembly-datum}, the passage
to a Wightman theory is the reconstruction stage following the RG estimates.
The distributional estimates give tempered limiting Schwinger functions.
The covariance, symmetry, Hermiticity, and normalization identities pass to
the limit because they are continuous linear identities on the test-function
spaces.  The reflection-positive Gram forms pass to the limit by the closed
cone argument used in
Proposition~\ref{prop:rg-reconstruction-estimates-give-limits}, provided the
directed finite-family estimates have actually been supplied.  The
positive-time semigroup regularity and the corrected OS-II growth package are
then exactly the remaining hypotheses of
Theorem~\ref{thm:os-reconstruction}.  Applying that theorem constructs the
Hilbert space, vacuum vector, translation representation, and operator-valued
tempered distribution whose Euclidean boundary values reproduce the
noncoincident limiting Schwinger functions.

These reconstruction requirements contain more information than the statement that
\((g_k,K_k)\) approaches a fixed point.  Finite-regulator reflection
positivity plus convergence of a few correlators supplies only the finite
windows under control.  OS reconstruction asks for the whole directed
positive-time Gram family, the positive-time translation semigroup, and the
OS-II analytic growth estimates.  These are the
model-specific burdens that remain for the ordinary short-range scalar
Wilson--Fisher target.

\begin{remark}[Current theorem status for the ordinary scalar target]
\label{rem:ordinary-scalar-target-status}
For massive superrenormalizable scalar theories, constructive methods can
produce Euclidean measures and hence Wightman theories after OS
reconstruction.  For an ordinary short-range three-dimensional critical
scalar model, a theorem at the Wilsonian fixed-point level would have to
construct Definition~\ref{def:short-range-scalar-block-spin-structure}, prove a
Banach-space fixed point or stable trajectory for the corresponding
short-range block-spin RG, establish
\eqref{eq:short-range-scalar-reconstruction-bound} for local fields and
their source extensions, prove reflection-compatible blocking as in
Lemma~\ref{lem:reflection-positive-block-spin-pullback} or an equivalent
positivity-preserving construction, and then supply the full
OS reconstruction requirements of
Definition~\ref{def:short-range-scalar-os-assembly-datum}.  This is stronger
than identifying a
candidate fixed point in an epsilon expansion, in a hierarchical model, or
in a finite projection truncation.
\end{remark}

\paragraph{The ordinary short-range scalar RG-to-OS assembly package.}
The preceding estimates can be organized into one theorem-facing package.  This
is not an additional theorem about the three-dimensional Ising or
Wilson--Fisher target; it is the precise list of estimates a model-specific
proof must supply before the OS reconstruction theorem can be invoked.

\begin{definition}[Short-range scalar RG-to-OS assembly package]
\label{def:short-range-rg-to-os-assembly-datum}
Fix a short-range scalar block-spin structure and source normalization as in
Definitions~\ref{def:short-range-scalar-block-spin-structure} and
\ref{def:short-range-scalar-os-assembly-datum}.  An RG-to-OS assembly package
for a critical short-range scalar target consists of the following
scale-indexed estimates.
\begin{enumerate}
  \item A tuned RG trajectory \((g_k,K_k)\) in a declared Banach polymer chart
  and a limiting source-extended chart
  \((C_{\ast,E},g_{\ast,E},K_{\ast,E},r_{\ast,E})\) for every finite test
  space \(E\), all defined on a fixed source polydisc \(P_{\rho_E}\).
  \item For every \(E\), a source-window error budget
  \begin{equation}
  \begin{aligned}
    \epsilon^{\rm src}_{E,k}
    =
    &\ \delta^C_{E,k}
      +L_E(\delta^g_{E,k}+\delta^K_{E,k})
      +\delta^r_{E,k}                                      \\
    &\ +A_E\Xi_E(R_{E,k})
      +\sigma_{E,k},
  \end{aligned}
  \label{eq:short-range-rg-to-os-source-budget}
  \end{equation}
  with \(\epsilon^{\rm src}_{E,k}\to0\).  The first line is the RG-chart
  error of \eqref{eq:short-range-source-window-chart-to-holomorphic-bound};
  \(A_E\Xi_E(R_{E,k})\) is the thermodynamic boundary tail of
  \eqref{eq:short-range-source-window-volume-tail}; and
  \(\sigma_{E,k}\) is the remaining cofinal-window, sampling, or
  finite-step error at scale \(k\).
  \item Uniform distribution bounds: for each \(n\) there are \(r_n\) and
  \(C_n\), independent of \(k\) and of the finite window \(E\), such that the
  \(n\)-source derivatives extracted from \(W_{k,E}\) satisfy
  \eqref{eq:short-range-source-window-uniform-schwartz-bound}.  Consequently,
  for every multi-index \(\beta\) in the \(E\)-window,
  \begin{equation}
    \left|
      \partial_z^\beta W_{k,E}(0)-\partial_z^\beta W_{\ast,E}(0)
    \right|
    \le
    \frac{\beta!}{\rho_E^\beta}\,
    \epsilon^{\rm src}_{E,k}.
    \label{eq:short-range-rg-to-os-cumulant-budget}
  \end{equation}
  \item Cofinal compatibility: the controlled finite test spaces are cofinal
  in the Schwartz test space, their restriction maps commute with the
  source-window functionals, and the fixed-window Cauchy estimates have the
  form required by Lemma~\ref{lem:rg-cofinal-finite-window-assembly}.
  \item Directed OS positivity and time-regularity: the positive-time Gram
  windows satisfy the directed positivity assembly hypotheses of
  Lemma~\ref{lem:rg-directed-os-positive-form-assembly}, the quotient
  stability and Gram-continuity estimates of
  Lemma~\ref{lem:rg-positive-time-quotient-semigroup-criterion}, and the
  corrected OS-II growth package in
  Definition~\ref{def:short-range-scalar-os-assembly-datum}.
\end{enumerate}
\end{definition}

When these estimates are supplied by a specific short-range scalar model, the
logic is fixed.  Equation~\eqref{eq:short-range-rg-to-os-cumulant-budget}
gives convergence of every finite source window.  Cofinal compatibility and
the uniform Schwartz bounds produce limiting tempered Schwinger
distributions.  Directed positivity, quotient-stable time translations, and
OS-II growth provide the hypotheses of
Theorem~\ref{thm:os-reconstruction}.  The conclusion is then a Wightman QFT
whose Euclidean Schwinger functions are the reconstructed limits.  The
unproved part for the ordinary critical scalar target is the model-specific
construction of the estimates in
Definition~\ref{def:short-range-rg-to-os-assembly-datum}; the implication
from those estimates to reconstruction is the functional-analytic assembly
just described.

\section{Transfer Estimates From Auxiliary Fixed-Point Models}
\label{sec:auxiliary-rg-transfer-estimates}

The theorem-level fixed points known in hierarchical scalar models,
fractional or long-range models, fermionic small-parameter models, and
tensor-network RG are not automatically fixed points of the ordinary
short-range scalar block-spin problem of
Definition~\ref{def:short-range-scalar-block-spin-structure}.  Their use for the
ordinary target is controlled by an auxiliary-to-target comparison, not by a
shared list of exponents.  This section records the estimate that such a
comparison must contain.

\begin{definition}[Auxiliary-to-short-range RG transfer estimates]
\label{def:auxiliary-to-short-range-rg-transfer-estimates}
Let
\[
  (\mathcal B_{\rm aux},\mathcal R_{\rm aux},V_{\ast}^{\rm aux})
  \qquad\text{and}\qquad
  (\mathcal B_{\rm sr},\mathcal R_{\rm sr},V_{\ast}^{\rm sr})
\]
be an auxiliary RG system and an ordinary short-range target RG system, with
Banach spaces, domains, fixed points, stable neighborhoods, source
extensions, and reconstruction maps already specified.  Transfer estimates from
the auxiliary system to the short-range target consist of:
\begin{enumerate}
  \item a map
  \(\Phi:\mathcal U_{\rm aux}\to\mathcal B_{\rm sr}\), defined on an
  auxiliary neighborhood that contains the orbit under consideration;
  \item a map \(\Psi:\mathcal O_{\rm aux}\to\mathcal O_{\rm sr}\) between the
  corresponding locally convex observable-data spaces;
  \item one-step RG defects
  \begin{equation}
    d_j(y_0)
    =
    \mathcal R_{\rm sr}\!\left(\Phi(y_j)\right)
    -
    \Phi(y_{j+1}),
    \qquad
    y_j=\mathcal R_{\rm aux}^{\,j}y_0 ,
    \label{eq:auxiliary-short-range-one-step-defect}
  \end{equation}
  with explicit norm bounds
  \(\|d_j(y_0)\|_{\mathcal B_{\rm sr}}\leq\delta_j(y_0)\);
  \item Lipschitz constants \(M_j\) for \(\mathcal R_{\rm sr}\) on a tube in
  \(\mathcal B_{\rm sr}\) containing both the true short-range orbit
  \(\mathcal R_{\rm sr}^{\,j}\Phi(y_0)\) and the transferred auxiliary orbit
  \(\Phi(y_j)\);
  \item comparison of tuning maps for the unstable coordinates,
  \begin{equation}
    \tau_{\rm sr}(\Phi y)
    =
    A\,\tau_{\rm aux}(y)+r(y),
    \label{eq:auxiliary-short-range-tuning-map}
  \end{equation}
  where \(A\) is a specified linear map between unstable coordinate spaces and
  \(r(y)\) is bounded in the target unstable norm;
  \item observable comparison estimates
  \begin{equation}
    p\!\left(
      \Psi\,\operatorname{Rec}^{\rm aux}_{n,k}(y_0)
      -
      \operatorname{Rec}^{\rm sr}_{n,k}(\Phi y_0)
    \right)
    \leq
    \varepsilon_{n,p,k}(y_0)
    \label{eq:auxiliary-short-range-observable-defect}
  \end{equation}
  for every declared observable seminorm \(p\) relevant to the claimed
  continuum statement.
\end{enumerate}
\end{definition}

The transfer estimates refine
Definition~\ref{def:comparison-of-rg-descriptions} in the direction needed
for ordinary short-range critical models.  The finite RG-orbit estimate is
straightforward, but it is the load-bearing estimate.  Let
\[
  x_j=\mathcal R_{\rm sr}^{\,j}\Phi(y_0),
  \qquad
  z_j=x_j-\Phi(y_j).
\]
Then \(z_0=0\) and
\begin{equation}
  z_{j+1}
  =
  \mathcal R_{\rm sr}(x_j)
  -
  \mathcal R_{\rm sr}(\Phi(y_j))
  +
  d_j(y_0).
  \label{eq:auxiliary-short-range-z-recursion}
\end{equation}
The Lipschitz estimate on the tube gives
\begin{equation}
  \|z_{j+1}\|_{\mathcal B_{\rm sr}}
  \leq
  M_j\|z_j\|_{\mathcal B_{\rm sr}}+\delta_j(y_0).
  \label{eq:auxiliary-short-range-z-bound-step}
\end{equation}
Iterating this scalar inequality yields
\begin{equation}
  \|z_n\|_{\mathcal B_{\rm sr}}
  \leq
  \sum_{j=0}^{n-1}
  \left(\prod_{\ell=j+1}^{n-1}M_\ell\right)\delta_j(y_0).
  \label{eq:auxiliary-short-range-telescoping-bound}
\end{equation}
Thus an auxiliary fixed-point theorem becomes a controlled statement about
the short-range target only when the one-step defects are summable after the
target RG amplification factors are included.

On a stable subspace, suppose \(M_j\leq\theta<1\) and
\(\delta_j(y_0)\leq Cq^j\) with \(0<q<1\).  Then
\begin{equation}
  \|z_n\|
  \leq
  C\sum_{j=0}^{n-1}\theta^{\,n-1-j}q^j
  =
  \begin{cases}
    C\dfrac{\theta^n-q^n}{\theta-q}, & \theta\neq q,\\[1.2em]
    Cn\theta^{\,n-1}, & \theta=q .
  \end{cases}
  \label{eq:auxiliary-short-range-stable-geometric-bound}
\end{equation}
Both terms tend to zero when \(\theta,q<1\).  This is the stable-direction
part of the comparison problem.

Relevant directions have the opposite behavior.  If a target unstable
coordinate has linear eigenvalue \(\lambda_a=L^{y_a}\), \(y_a>0\), and
one-step comparison errors \(e_{a,j}\), then the finite-depth bound is
\begin{equation}
  |u_{a,N}|
  \leq
  |\lambda_a|^N |u_{a,0}|
  +
  \sum_{j=0}^{N-1}|\lambda_a|^{\,N-1-j}|e_{a,j}|.
  \label{eq:auxiliary-short-range-relevant-amplification}
\end{equation}
Consequently, if \(y\) lies on the auxiliary critical surface
\(\tau_{\rm aux}(y)=0\), the target initial point \(\Phi y\) lies on the
target critical surface only when the remainder \(r(y)\) in
\eqref{eq:auxiliary-short-range-tuning-map} is itself tuned away, or is small
to the finite-depth accuracy demanded by
\eqref{eq:auxiliary-short-range-relevant-amplification}.  This is why a
comparison between a hierarchical or long-range theorem and an ordinary
short-range scalar theorem must include the unstable tuning map, not only the
stable contraction estimate.

Finally, the observable estimate
\eqref{eq:auxiliary-short-range-observable-defect} must be combined with
\eqref{eq:auxiliary-short-range-telescoping-bound}.  If
\(\operatorname{Rec}^{\rm sr}_{n,k}\) is locally Lipschitz in the RG chart
with respect to a seminorm \(p\), then the observable error is bounded by the
explicit auxiliary-observable defect plus the Lipschitz image of the orbit
error \(z_k\).  Only after both terms vanish in the chosen observable
topology does an auxiliary theorem transfer to the ordinary short-range QFT
claim.  Without this transfer structure, hierarchical, long-range, tensor, and
functional-RG fixed points should be read as theorem-level benchmarks or
computational evidence in their own frameworks, not as a proof of the
ordinary short-range Wilson--Fisher QFT.

\paragraph{Transfer to a projective observable germ.}
The comparison must be made window by window, because a continuum local-QFT
claim is a projective statement about all reconstruction windows.  Fix a
finite observation window \(\alpha\) in the directed family of the short-range
target.  Suppose the auxiliary theorem supplies a corresponding auxiliary
window and an auxiliary limiting observable \(O^{\rm aux}_{\ast,\alpha}\) such
that, after its normalization,
\begin{equation}
  p_\alpha^{\rm sr}\!\left(
    \Psi\,\operatorname{Rec}^{\rm aux}_{\alpha,k}(y_0)
    -
    \Psi O^{\rm aux}_{\ast,\alpha}
  \right)
  \leq
  a_{\alpha,k}.
  \label{eq:auxiliary-window-convergence-error}
\end{equation}
Assume further that the comparison of observable reconstructions has the
finite-window defect
\begin{equation}
  p_\alpha^{\rm sr}\!\left(
    \operatorname{Rec}^{\rm sr}_{\alpha,k}(\Phi y_0)
    -
    \Psi\,\operatorname{Rec}^{\rm aux}_{\alpha,k}(y_0)
  \right)
  \leq
  \varepsilon_{\alpha,k},
  \label{eq:auxiliary-target-observable-window-defect}
\end{equation}
and that the short-range reconstruction map is Lipschitz on the target tube:
\begin{equation}
  p_\alpha^{\rm sr}\!\left(
    \operatorname{Rec}^{\rm sr}_{\alpha,k}(x_0)
    -
    \operatorname{Rec}^{\rm sr}_{\alpha,k}(\Phi y_0)
  \right)
  \leq
  L_{\alpha,k}\|x_k-\Phi(y_k)\|_{\mathcal B_{\rm sr}} .
  \label{eq:short-range-window-lipschitz-orbit-error}
\end{equation}
Here \(x_0\) is the actual short-range initial point whose window is being
claimed; in the simplest transferred construction \(x_0=\Phi y_0\), but the
separate notation is useful when the target microscopic regulator is only
close to the transferred auxiliary initial condition.  If the allowed
normalization of the short-range window differs from the pushed-forward
auxiliary normalization by at most \(n_{\alpha,k}\), the triangle inequality
and \eqref{eq:auxiliary-short-range-telescoping-bound} give the finite
projective-window transfer estimate
\begin{equation}
\begin{aligned}
  p_\alpha^{\rm sr}\!\left(
    g_{\rm sr}\operatorname{Rec}^{\rm sr}_{\alpha,k}(x_0)
    -
    \Psi O^{\rm aux}_{\ast,\alpha}
  \right)
  &\leq
  n_{\alpha,k}
  +a_{\alpha,k}
  +\varepsilon_{\alpha,k}                                      \\
  &\quad+
  L_{\alpha,k}
  \sum_{j=0}^{k-1}
  \left(\prod_{\ell=j+1}^{k-1}M_\ell\right)\delta_j(y_0).
\end{aligned}
  \label{eq:auxiliary-target-projective-window-estimate}
\end{equation}
Thus an auxiliary fixed point transfers to the ordinary short-range
observable germ only after the right-hand side of
\eqref{eq:auxiliary-target-projective-window-estimate} tends to zero for
every window \(\alpha\) in the directed family.  The estimate separates four
different analytic tasks: convergence of the auxiliary observable window,
comparison of auxiliary and target reconstructions, control of the target
orbit defect, and compatibility of the normalization group.  A shared
critical exponent controls at most one contribution to one family of
windows; it does not provide the projective observable-germ transfer.

\section{Source-Extended RG, Scaling Operators, And Contact Terms}

A fixed interaction is not yet a QFT spectrum.  Scaling operators are
obtained by extending the RG map to sources and by reconstructing the
resulting functional derivatives as distributions.  This section records the
minimal theorem-level structure, because otherwise the phrase ``operator
dimension'' may mean only an eigenvalue of a finite list of couplings.

Let \(A\) be a set of source labels.  For each \(\alpha\in A\), let
\(\mathcal J_\alpha\) be a locally convex space of test sources on
\(\mathbb R^D\), for instance \(C_c^\infty(\mathbb R^D)\) or
\(\mathcal S(\mathbb R^D)\).  A regulated source-dependent generating
functional has the form
\begin{equation}
  Z_k(V;J)
  =
  \int
  \exp\!\left(
    -S_k(\phi;V)
    +\sum_{\alpha\in A}\langle J_\alpha,\mathcal O_{\alpha,k}(\phi)\rangle
  \right)\,\dd\mu_k(\phi),
  \label{eq:source-extended-rg-generating-functional}
\end{equation}
where \(\langle J_\alpha,\mathcal O_{\alpha,k}\rangle\) denotes the pairing
between a test source and a regulated insertion.  In a gauge theory, the
source direction must either couple to a gauge-invariant lattice observable,
to a Wilson-line or Wilson-loop observable with its path data included, or
to a class in the BV observable cohomology.  A source coupled to a
gauge-fixed colored field variable is not, by itself, a nonperturbative
observable.

\begin{definition}[Source-extended RG datum]
\label{def:source-extended-rg-datum}
A source-extended RG datum near a fixed point \(V_\ast\) consists of:
\begin{enumerate}
  \item a Banach RG chart \(\mathcal B\), an RG map
  \(\mathcal R:\mathcal U\to\mathcal B\), and a fixed point
  \(\mathcal R(V_\ast)=V_\ast\);
  \item a locally convex source space
  \(\mathcal J=\bigoplus_{\alpha\in A}\mathcal J_\alpha\);
  \item a map
  \begin{equation}
    \widehat{\mathcal R}:\mathcal U\times\mathcal V_J
    \longrightarrow \mathcal B\times\mathcal J,
    \qquad
    \widehat{\mathcal R}(V,J)=
    \bigl(\mathcal R(V),\,\mathcal M_VJ+\mathcal N_V(J)\bigr),
    \label{eq:source-extended-rg-map}
  \end{equation}
  where \(\mathcal V_J\) is a neighborhood of \(0\in\mathcal J\),
  \(\mathcal M_V\) is linear in \(J\), and
  \(\mathcal N_V(J)=O(J^2)\) in the declared source topology;
  \item an exact finite-regulator identity
  \begin{equation}
    Z_k(V;J)
    =
    e^{A_k(V;J)}
    Z_{k-1}\!\left(\widehat{\mathcal R}(V,J)\right),
    \label{eq:source-rg-exact-identity}
  \end{equation}
  with \(A_k(V;J)\) a local normalization functional in the sources;
  \item reconstruction maps taking \(n\) source derivatives of
  \(\log Z_k(V;J)\) at \(J=0\) to connected distributions on
  \((\mathbb R^D)^n\), or to distributions on configuration space away from
  collision diagonals.
\end{enumerate}
The local functional \(A_k(V;J)\) is part of the datum.  Its source
derivatives are precisely where contact terms generated by the RG step are
recorded.
\end{definition}

At a fixed point, the linear source map
\begin{equation}
  \mathcal M_\ast=\mathcal M_{V_\ast}
  =
  D_J\widehat{\mathcal R}(V_\ast,0)
  \label{eq:linear-source-rg-map}
\end{equation}
contains the scaling data.  If a source \(J_\alpha\) couples to a local field
whose scaling dimension is \(\Delta_\alpha\), then the constant-source
coordinate has RG exponent
\begin{equation}
  y_\alpha=D-\Delta_\alpha .
  \label{eq:source-exponent-operator-dimension}
\end{equation}
This relation is fixed by the invariant pairing
\(\int J_\alpha(x)\mathcal O_\alpha(x)\,\dd^Dx\): under
\(x=Lx'\), a field with
\(\mathcal O_\alpha(Lx')=L^{-\Delta_\alpha}\mathcal O_\alpha(x')\) requires
\begin{equation}
  J'_\alpha(x')=L^{D-\Delta_\alpha}J_\alpha(Lx').
  \label{eq:source-rescaling-from-pairing}
\end{equation}
Thus the dimension is not assigned to a formal monomial alone; it is read
from the source direction after redundant directions, null fields, and
contact redefinitions have been quotiented as appropriate.

\begin{definition}[Scaling source and scaling field]
\label{def:scaling-source-and-field}
Let \(\mathcal M_\ast\) be the linearized source RG map.  A scaling source
direction is a nonzero equivalence class \([j_\alpha]\) in the source space
modulo null source directions and source redefinitions whose functional
derivatives vanish at separated points, such that
\begin{equation}
  \mathcal M_\ast [j_\alpha]
  =
  L^{D-\Delta_\alpha}[j_\alpha].
  \label{eq:scaling-source-eigenvalue}
\end{equation}
The corresponding scaling field is the distribution-valued functional
derivative of the connected generating functional in the direction
\([j_\alpha]\), reconstructed in the continuum limit.  In a BV formulation,
the equivalence class is taken in \(Q_{S_\ast}\)-cohomology of observables;
in a lattice gauge formulation, it is taken among gauge-invariant source
directions.
\end{definition}

The quotient in Definition~\ref{def:scaling-source-and-field} is not a
matter of taste.  If two source directions differ by a local source
counterterm, their separated-point connected correlators agree while their
extensions across coincident points may differ.  The following definition
fixes the vocabulary used in the rest of the monograph.

\begin{definition}[Contact term]
\label{def:rg-contact-term}
Let \(n\geq2\), and let
\begin{equation}
  \mathfrak D_n
  =
  \bigcup_{1\leq i<j\leq n}
  \{(x_1,\ldots,x_n)\in(\mathbb R^D)^n:\ x_i=x_j\}
  \label{eq:collision-diagonal-union}
\end{equation}
be the union of pairwise collision diagonals.  A contact term in an
\(n\)-point distribution is a distribution
\[
  C\in\mathcal D'((\mathbb R^D)^n)
  \]
whose support is contained in \(\mathfrak D_n\).  Equivalently,
\(\langle C,f\rangle=0\) for every
\(f\in C_c^\infty((\mathbb R^D)^n)\) whose support is disjoint from
\(\mathfrak D_n\).  Contact terms are not ordinary separated-point
functions.  In a source formalism they are generated by local functionals of
the sources and their derivatives.
\end{definition}

Let
\begin{equation}
  \operatorname{Conf}_n(\mathbb R^D)
  =
  (\mathbb R^D)^n\setminus\mathfrak D_n
  \label{eq:configuration-space-distinct-points}
\end{equation}
be the configuration space of ordered distinct points.  For
\(f\in C_c^\infty(\operatorname{Conf}_n(\mathbb R^D))\), define the dilation
of the test function by
\begin{equation}
  (S_Lf)(x_1,\ldots,x_n)
  =
  L^{-nD}
  f(L^{-1}x_1,\ldots,L^{-1}x_n).
  \label{eq:test-function-dilation-for-correlators}
\end{equation}
This convention is chosen so that
\begin{equation}
  \langle G,S_Lf\rangle
  =
  \int G(Ly_1,\ldots,Ly_n)f(y_1,\ldots,y_n)\,
       \dd^Dy_1\cdots\dd^Dy_n
  \label{eq:test-dilation-change-of-variables}
\end{equation}
whenever \(G\) is represented by an integrable function on the support of
the test function.

\paragraph{Separated-point scaling from source-extended RG.}
The nontrivial step in a source-extended RG construction is proving the
finite-scale identity below for actual source derivatives of the regulated
theory.  Once that identity is available with a vanishing error, its
separated-point consequence is a distributional limit statement.  Fix labels
\(\alpha_1,\ldots,\alpha_n\) with source eigenvalues
\(L^{D-\Delta_{\alpha_i}}\).  Suppose that the source-extended RG datum gives
connected finite-regulator distributions
\[
  G_{k;\alpha_1\ldots\alpha_n}
  \in
  \mathcal D'(\operatorname{Conf}_n(\mathbb R^D))
  \]
which converge to
\[
  G_{\ast;\alpha_1\ldots\alpha_n}
  \in
  \mathcal D'(\operatorname{Conf}_n(\mathbb R^D))
  \]
in the usual distribution topology on every compact subset of
\(\operatorname{Conf}_n(\mathbb R^D)\).  Assume moreover that for every
\(f\in C_c^\infty(\operatorname{Conf}_n(\mathbb R^D))\) the finite-scale RG
identity implies
\begin{equation}
  \left|
  \langle G_{k;\alpha_1\ldots\alpha_n},S_Lf\rangle
  -
  L^{-\sum_i\Delta_{\alpha_i}}
  \langle G_{k-1;\alpha_1\ldots\alpha_n},f\rangle
  \right|
  \leq \varepsilon_k(f),
  \label{eq:finite-scale-separated-source-scaling}
\end{equation}
with \(\varepsilon_k(f)\to0\).  Then the limiting separated-point
distribution is homogeneous under scale \(L\):
\begin{equation}
  \langle G_{\ast;\alpha_1\ldots\alpha_n},S_Lf\rangle
  =
  L^{-\sum_i\Delta_{\alpha_i}}
  \langle G_{\ast;\alpha_1\ldots\alpha_n},f\rangle
  \label{eq:limiting-separated-source-scaling}
\end{equation}
for every \(f\in C_c^\infty(\operatorname{Conf}_n(\mathbb R^D))\).

If \(T_{\ast;\alpha_1\ldots\alpha_n}\in
\mathcal D'((\mathbb R^D)^n)\) is any extension of
\(G_{\ast;\alpha_1\ldots\alpha_n}\) to all of \((\mathbb R^D)^n\), then the
distribution
\begin{equation}
  H_L(f)
  =
  \langle T_{\ast;\alpha_1\ldots\alpha_n},S_Lf\rangle
  -
  L^{-\sum_i\Delta_{\alpha_i}}
  \langle T_{\ast;\alpha_1\ldots\alpha_n},f\rangle
  \label{eq:scaling-violation-contact-distribution}
\end{equation}
is a contact term.

The verification is a support calculation for distributions, not a separate
RG theorem.
The dilation \(S_L\) maps
\(C_c^\infty(\operatorname{Conf}_n(\mathbb R^D))\) continuously to itself.
By distributional convergence on configuration space,
\[
  \langle G_{k;\alpha_1\ldots\alpha_n},S_Lf\rangle
  \longrightarrow
  \langle G_{\ast;\alpha_1\ldots\alpha_n},S_Lf\rangle
\]
and
\[
  \langle G_{k-1;\alpha_1\ldots\alpha_n},f\rangle
  \longrightarrow
  \langle G_{\ast;\alpha_1\ldots\alpha_n},f\rangle .
\]
Taking \(k\to\infty\) in
\eqref{eq:finite-scale-separated-source-scaling} gives
\eqref{eq:limiting-separated-source-scaling}, because
\(\varepsilon_k(f)\to0\).

For the last assertion, let \(f\in
C_c^\infty((\mathbb R^D)^n)\) have support disjoint from
\(\mathfrak D_n\).  Then \(f\) and \(S_Lf\) are test functions on
\(\operatorname{Conf}_n(\mathbb R^D)\), and the extension
\(T_{\ast;\alpha_1\ldots\alpha_n}\) agrees with
\(G_{\ast;\alpha_1\ldots\alpha_n}\) on both of them.  Therefore
\eqref{eq:limiting-separated-source-scaling} gives \(H_L(f)=0\).  A
distribution that vanishes on every test function supported in the open set
\((\mathbb R^D)^n\setminus\mathfrak D_n\) has support contained in
\(\mathfrak D_n\), by the definition of support of a distribution.  Hence
\(H_L\) is a contact term in the sense of
Definition~\ref{def:rg-contact-term}.

This is the logical order.  Source eigenvalues determine
candidate scaling dimensions.  Reconstruction estimates make the source
derivatives into distributions.  The finite-scale RG identity then proves
homogeneity at separated points.  Extension across collision diagonals is an
additional renormalization problem; different extensions differ by contact
terms, and local source counterterms change precisely those contact terms.
This is why anomalous dimensions, OPE coefficients at separated points, and
contact terms must be kept as distinct pieces of data.

\begin{remark}[Status of short-range scalar and gauge targets]
\label{rem:short-range-scalar-gauge-rg-status}
Constructive massive \(\phi^4_3\) and related superrenormalizable models give
Wightman QFTs after their Euclidean constructions and OS reconstruction are
completed.  That achievement is logically distinct from constructing the
critical Wilson--Fisher CFT as a non-Gaussian fixed point of a short-range
block-spin map with stable manifold, operator spectrum, and reconstruction
estimates of the form
\eqref{eq:rg-reconstruction-estimate}.  Hierarchical scalar models,
fractional long-range models, fermionic small-parameter models, and rigorous
tensor RG supply theorem-level benchmarks for parts of this structure.

For gauge theories the corresponding theorem must be formulated through
Definition~\ref{def:gauge-compatible-wilsonian-rg-structure}.  A lattice proof
would construct the RG map on gauge-invariant measures and Wilson-line
observables with reflection-positivity control.  A continuum perturbative
proof would construct the same flow in a regularized BV complex, prove the
quantum master equation after each RG step, and identify the physical
observables as cohomology classes before reconstruction.  These are stronger
requirements than the computation of beta functions for a finite list of
local coefficients.

In four-dimensional scalar \(\phi^4\), perturbation theory supplies a formal
renormalized coordinate system.  Existing triviality results and expectations
belong to particular regulator classes and limiting procedures.  The
existence of a UV-complete nonperturbative QFT whose short-distance
expansion agrees with four-dimensional \(\phi^4\) perturbation theory to all
orders remains outside the theorem-level results recorded here.
\end{remark}

\section{Relation To Universality Classes}

In this monograph, the phrase \emph{universality class} is not used as a
synonym for ``same symmetries and same dimension.''  It denotes a convergence
statement in a declared topology.  The minimal statement is the following.

\begin{definition}[Wilsonian universality statement]
\label{def:wilsonian-universality-statement}
A Wilsonian universality statement consists of:
\begin{enumerate}
  \item an index set \(I\) of microscopic regulators or bare finite-cutoff
  models;
  \item for each \(i\in I\), a tuned initial point \(V_i\) in a common Banach
  RG chart \((\mathcal B,\|\cdot\|)\), or in charts identified by specified
  coordinate maps;
  \item an RG map \(\mathcal R:\mathcal U\to\mathcal B\), a neighborhood
  \(\mathcal N\subset\mathcal U\), and a fixed point \(V_\ast\in\mathcal N\);
  \item a topological vector space \(\mathcal O\) of normalized long-distance
  observable data, for example finite lists of smeared Schwinger functions in
  a distribution topology;
  \item reconstruction maps
  \[
    \operatorname{Rec}_{i,n}:\mathcal U\to\mathcal O
  \]
  giving the \(n\)-step long-distance observable data of the microscopic
  model \(i\);
  \item a normalization group \(G\) acting continuously on \(\mathcal O\),
  encoding allowed metric factors, field normalizations, and coordinate
  rescalings.
\end{enumerate}
This is a theorem-level universality statement if, for every \(i\in I\),
there are \(g_i\in G\) and \(O_\ast\in\mathcal O\) such that
\[
  \mathcal R^n(V_i)\in\mathcal N\quad(n\gg1),\qquad
  \mathcal R^n(V_i)\longrightarrow V_\ast\quad\text{in }\mathcal B,
\]
and
\[
  g_i\,\operatorname{Rec}_{i,n}(V_i)
  \longrightarrow O_\ast
  \qquad\text{in }\mathcal O .
\]
The common limit \(O_\ast\) is part of the assertion; without the
reconstruction convergence, convergence of local couplings alone is not a
QFT universality theorem.
\end{definition}

\paragraph{Observable-germ formulation.}
For an actual QFT the reconstructed observable is rarely a single Banach-space
coordinate.  It is normally a compatible family of finite tests: finitely
many smeared \(n\)-point functions, finite lists of Wilson loops, finite
momentum windows, finite source derivatives, or finite spectral windows.
Thus the topology in
Definition~\ref{def:wilsonian-universality-statement} is most naturally
given by a projective system.

\begin{definition}[Observable-germ universality class]
\label{def:observable-germ-universality-class}
An observable-germ universality structure is a Wilsonian universality statement
together with the following refinement of \(\mathcal O\).
\begin{enumerate}
  \item A directed set \(\mathfrak A\) of finite observation windows.
  \item For each \(\alpha\in\mathfrak A\), a Banach space
  \(\mathcal O_\alpha\) with norm \(p_\alpha\), and for
  \(\alpha\preceq\beta\) a continuous projection
  \(\pi_{\alpha\beta}:\mathcal O_\beta\to\mathcal O_\alpha\) satisfying
  \[
    \pi_{\alpha\gamma}
    =
    \pi_{\alpha\beta}\pi_{\beta\gamma},
    \qquad
    \alpha\preceq\beta\preceq\gamma .
  \]
  \item The projective-limit observable space
  \[
    \mathcal O_{\mathrm{germ}}
    =
    \varprojlim_{\alpha\in\mathfrak A}\mathcal O_\alpha
    =
    \left\{
      (O_\alpha)_\alpha:\pi_{\alpha\beta}O_\beta=O_\alpha
      \text{ for }\alpha\preceq\beta
    \right\},
  \]
  equipped with the seminorms \(p_\alpha(O)=p_\alpha(O_\alpha)\).
  \item Reconstruction maps
  \[
    \operatorname{Rec}_{i,n}^{\alpha}:
    \mathcal U\longrightarrow\mathcal O_\alpha
  \]
  compatible with the projections:
  \[
    \pi_{\alpha\beta}\operatorname{Rec}_{i,n}^{\beta}
    =
    \operatorname{Rec}_{i,n}^{\alpha}.
  \]
  \item A normalization group \(G\) acting by compatible continuous maps on
  all \(\mathcal O_\alpha\).
\end{enumerate}
The microscopic regulator \(i\) belongs to the observable-germ universality
class represented by the normalized germ \(O_\ast\in\mathcal O_{\mathrm{germ}}\)
if there exists \(g_i\in G\) such that, for every finite observation window
\(\alpha\),
\begin{equation}
  p_\alpha\!\left(
    g_i\,\operatorname{Rec}_{i,n}^{\alpha}(V_i)
    -
    (O_\ast)_\alpha
  \right)
  \longrightarrow0 .
  \label{eq:observable-germ-universality-convergence}
\end{equation}
The word \emph{class} therefore refers to an equivalence class of normalized
observable germs in this projective topology.  Agreement of finitely many
critical exponents, amplitude ratios, or correlation functions is a
finite-window statement until the remaining windows are controlled.
\end{definition}

The preceding definition deliberately uses observable windows rather than
the phrase ``the same theory.''  Equality of a few numerical quantities, or
even of a large but finite list of reconstructed correlators, is not equality
of local QFT data.  The bridge to a local quantum field theory requires that
the projective family contain enough test-function windows, positivity
conditions, covariance identities, and growth bounds to feed one of the
reconstruction frameworks developed in
Chapters~\ref{ch:volume-iv-osterwalder-schrader-reconstruction} and
\ref{ch:volume-iv-wightman-fields-reconstruction}.

\begin{definition}[Local-QFT-strength observable germ]
\label{def:local-qft-strength-observable-germ}
An observable-germ structure is \emph{local-QFT-strength} for an
Euclidean scalar or gauge-invariant observable sector if its directed family
\(\mathfrak A\) contains, as compatible subwindows, the following data.
\begin{enumerate}
  \item For each finite list of observable labels
  \(a_1,\ldots,a_n\) and each finite-dimensional subspace
  \(E\subset\mathcal S((\mathbb R^D)^n)\) of test functions, a window that
  records the values of a distribution
  \(S_{a_1\ldots a_n}\in\mathcal S'((\mathbb R^D)^n)\) on \(E\).  The
  projective limit over \(E\) recovers the full tempered distribution only
  after the compatibility and seminorm-bound conditions of
  Lemma~\ref{lem:rg-projective-window-extension-distribution} are supplied;
  finite pointwise or moment data alone are a smaller collection of tests.
  \item Windows recording the algebraic identities required of the sector:
  permutation or graded permutation covariance, Euclidean covariance,
  reflection compatibility for the chosen time direction, and the local
  source-counterterm subspace whose derivatives are distributions supported
  on collision diagonals.
  \item Windows recording the positivity forms used in reconstruction.  In
  the OS route these are the finite reflection-positivity Gram matrices for
  test functions supported at positive Euclidean time.  In a directly
  Lorentzian route they are the Wightman positivity forms on the polynomial
  test-field algebra.
  \item The growth or regularity hypotheses needed by the chosen
  reconstruction theorem.  For the OS-to-Wightman route this includes the
  corrected OS-II growth hypothesis used in
  Chapter~\ref{ch:volume-iv-osterwalder-schrader-reconstruction}; for the
  Wightman route it includes temperedness, the spectrum condition, locality,
  and the common invariant domain structure.
  \item If vacuum uniqueness or particle data are part of the claim, windows
  for cluster properties or spectral projections in the declared observable
  topology.
\end{enumerate}
The normalization group is restricted to transformations of the observable
labels, coordinates, metric factors, and source normalizations that preserve
these identities and positivity forms.
\end{definition}

\paragraph{Projective distribution-window extension.}
The first item in
Definition~\ref{def:local-qft-strength-observable-germ} contains a functional
analytic obligation that is easy to hide if one speaks only of many finite
windows.  Fix \(n\) and write
\(X=(\mathbb R^D)^n\).  For an integer \(r\geq0\), set
\begin{equation}
  q_r(f)
  =
  \max_{|\alpha|\leq r}\sup_{x\in X}
  (1+|x|)^r\,|\partial^\alpha f(x)|,
  \qquad
  f\in\mathcal S(X),
  \label{eq:rg-schwartz-seminorm-window}
\end{equation}
where \(\alpha\) is a multi-index on \(X\).  These seminorms generate the
Schwartz topology up to the usual harmless replacement by an equivalent
countable family.

\begin{lemma}[Projective finite windows define a distribution]
\label{lem:rg-projective-window-extension-distribution}
Let \(\mathcal E\) be a directed family of finite-dimensional subspaces of
\(\mathcal S(X)\), ordered by inclusion, and suppose that
\(E_\infty=\bigcup_{E\in\mathcal E}E\) is dense in
\(\mathcal S(X)\).  For each \(E\in\mathcal E\), let
\(S_E:E\to\mathbb C\) be linear.  Assume:
\begin{enumerate}
  \item whenever \(E\subset F\), \(S_F|_E=S_E\);
  \item there exist \(r\in\mathbb Z_{\geq0}\) and \(C<\infty\) such that
  \begin{equation}
    |S_E(f)|\leq C\,q_r(f)
    \qquad
    (E\in\mathcal E,\ f\in E).
    \label{eq:rg-projective-window-uniform-bound}
  \end{equation}
\end{enumerate}
Then there is a unique tempered distribution
\(S\in\mathcal S'(X)\) satisfying \(S|_E=S_E\) for every
\(E\in\mathcal E\).  Moreover \(S\) obeys the same bound
\[
  |S(f)|\leq C\,q_r(f),
  \qquad f\in\mathcal S(X).
\]
\end{lemma}

\begin{proof}
Compatibility first produces a single algebraic functional on
\(E_\infty\).  If \(f\in E\cap F\), choose \(H\in\mathcal E\) with
\(E\subset H\) and \(F\subset H\).  Then
\[
  S_E(f)=S_H(f)=S_F(f),
\]
so \(S_\infty(f):=S_E(f)\) for \(f\in E\) is well defined on
\(E_\infty\).  Linearity follows after passing to a common upper window
containing the finitely many vectors being combined.  The bound
\eqref{eq:rg-projective-window-uniform-bound} gives
\[
  |S_\infty(f)|\leq C q_r(f),
  \qquad f\in E_\infty .
\]

Let \(f\in\mathcal S(X)\).  Choose \(f_j\in E_\infty\) with
\(f_j\to f\) in the Schwartz topology.  Then \(q_r(f_j-f_k)\to0\), hence
\[
  |S_\infty(f_j)-S_\infty(f_k)|
  \leq C q_r(f_j-f_k),
\]
so \(S_\infty(f_j)\) is Cauchy in \(\mathbb C\).  If \(g_j\in E_\infty\)
is another approximating sequence, then \(q_r(f_j-g_j)\to0\), so the two
limits agree.  Define
\[
  S(f)=\lim_{j\to\infty}S_\infty(f_j).
\]
The definition is linear by applying the same construction to approximating
sequences for finite linear combinations, and the estimate
\(|S(f)|\leq Cq_r(f)\) follows by taking the limit of
\(|S_\infty(f_j)|\leq Cq_r(f_j)\).  Thus \(S\) is continuous on
\(\mathcal S(X)\), hence \(S\in\mathcal S'(X)\), and it extends each
\(S_E\).  If \(T\in\mathcal S'(X)\) is another continuous extension, then
\(T-S\) vanishes on the dense subspace \(E_\infty\); continuity gives
\(T=S\).
\end{proof}

This lemma isolates the exact theorem-facing content of the phrase
``distribution window.''  Compatibility supplies a single algebraic
functional; the uniform Schwartz-seminorm bound supplies tempered
continuity.  In a genuine Wilsonian RG theorem the estimates must be produced
for the Schwinger hierarchy and for the source derivatives that generate the
desired local observables.  After this distributional object exists, the
finite reflection-positivity windows below test whether it is eligible for
OS reconstruction.

\paragraph{Cofinal assembly from regulated finite windows.}
The previous lemma starts from limiting window functionals.  An RG proof
usually produces regulated window functionals \(S_{k,N}\) at scale \(k\),
where \(N\) records how large a finite test window has been controlled.  The
following criterion is the diagonal step needed before the limiting
functionals in Lemma~\ref{lem:rg-projective-window-extension-distribution}
exist.

\begin{lemma}[Cofinal finite-window assembly]
\label{lem:rg-cofinal-finite-window-assembly}
Let
\[
  E_1\subset E_2\subset\cdots\subset\mathcal S(X)
\]
be finite-dimensional subspaces whose union is dense in \(\mathcal S(X)\).
Let \(N_k\to\infty\), and suppose that for every \(k\) and every
\(N\le N_k\) there is a linear functional
\[
  S_{k,N}:E_N\to\mathbb C .
\]
Assume:
\begin{enumerate}
  \item finite-scale compatibility:
  \(S_{k,M}|_{E_N}=S_{k,N}\) whenever \(N\le M\le N_k\);
  \item a uniform seminorm bound: there exist \(r\) and \(C\), independent of
  \(k\) and \(N\), such that
  \[
    |S_{k,N}(f)|\le C q_r(f),
    \qquad f\in E_N,\quad N\le N_k;
  \]
  \item cofinal Cauchy control: for each fixed \(N\),
  \begin{equation}
    \eta_{N,k}
    =
    \sup_{\ell\ge k}
    \sup_{\substack{f\in E_N\\ q_r(f)\le1}}
    |S_{k,N}(f)-S_{\ell,N}(f)|
    \longrightarrow0
    \label{eq:cofinal-window-cauchy-control}
  \end{equation}
  as \(k\to\infty\), with the supremum restricted to
  \(k,\ell\) for which \(N\le N_k,N_\ell\).
\end{enumerate}
Then there are compatible limiting functionals
\[
  S_N:E_N\to\mathbb C,\qquad S_M|_{E_N}=S_N\quad(N\le M),
\]
with \(|S_N(f)|\le Cq_r(f)\).  Hence
Lemma~\ref{lem:rg-projective-window-extension-distribution} produces a unique
tempered distribution \(S\in\mathcal S'(X)\) extending all \(S_N\).
Moreover one may choose a strictly increasing scale subsequence \(k_j\) with
\(N_{k_j}\ge j\) and
\begin{equation}
  \eta_{N,k_j}\le 2^{-j},
  \qquad 1\le N\le j .
  \label{eq:cofinal-diagonal-window-schedule}
\end{equation}
Along this subsequence every fixed finite window \(E_N\) converges to its
limit at a quantified rate after \(j\ge N\).
\end{lemma}

\begin{proof}
Fix \(N\).  Since \(N_k\to\infty\), the functionals \(S_{k,N}\) are defined
for all sufficiently large \(k\).  The Cauchy control
\eqref{eq:cofinal-window-cauchy-control} says that \(S_{k,N}\) is Cauchy in
the operator seminorm
\[
  \|T\|_{E_N,r}
  =
  \sup_{\substack{f\in E_N\\ q_r(f)\le1}} |T(f)|.
\]
The target space is finite dimensional, so this normed space of linear
functionals is complete.  Define \(S_N\) to be the limit of \(S_{k,N}\) in
that norm.  The bound \(|S_N(f)|\le Cq_r(f)\) follows by taking the limit of
the uniform bound for \(S_{k,N}\).

If \(N\le M\), choose \(k\) so large that \(M\le N_k\).  Finite-scale
compatibility gives
\[
  S_{k,M}|_{E_N}=S_{k,N}.
\]
Taking \(k\to\infty\) in the operator seminorms on \(E_M\) and \(E_N\) gives
\[
  S_M|_{E_N}=S_N .
\]
Thus the limiting windows satisfy the hypotheses of
Lemma~\ref{lem:rg-projective-window-extension-distribution}, and the
tempered distribution follows.

It remains to spell out the diagonal schedule.  For a fixed \(j\), the set
\(\{1,\ldots,j\}\) is finite.  Since each \(\eta_{N,k}\to0\), and since
\(N_k\to\infty\), there exists \(k_j\) larger than \(k_{j-1}\) such that
\[
  N_{k_j}\ge j,\qquad
  \eta_{N,k_j}\le2^{-j}\quad(1\le N\le j).
\]
Choosing these \(k_j\) recursively gives
\eqref{eq:cofinal-diagonal-window-schedule}.  For fixed \(N\) and
\(j\ge N\), the definition of \(\eta_{N,k_j}\) bounds the distance from
\(S_{k_j,N}\) to the limiting functional \(S_N\) by \(2^{-j}\) in the same
operator seminorm.
\end{proof}

This lemma identifies a common hidden gap in RG reconstruction arguments.
Controlling one finite window at each scale is insufficient: the number of
controlled windows must be cofinal, the already-controlled windows must be
Cauchy as the regulator is removed, and the seminorm bound must be uniform in
the window.  The lemma supplies the diagonal functional-analytic step.  It
does not supply the model-specific polymer or cluster estimates that prove
the hypotheses.

\paragraph{Finite OS-positivity Gram-matrix bounds.}
The positivity windows in
Definition~\ref{def:local-qft-strength-observable-germ} are not decorative.
They are finite Gram matrices whose limits must be controlled before
Osterwalder--Schrader reconstruction can be invoked.  Fix a finite family
\(\mathcal F=\{F_1,\ldots,F_m\}\) of polynomial Euclidean observable test
vectors supported at positive Euclidean time, and let \(\Theta\) be the
Euclidean time reflection with the conjugation appropriate to the observable
sector.  A regulated Schwinger hierarchy gives the finite matrix
\begin{equation}
  G_{k,\mathcal F;i j}
  =
  S_k\bigl((\Theta F_i)F_j\bigr),
  \qquad
  1\le i,j\le m .
  \label{eq:rg-finite-os-gram-matrix}
\end{equation}
Reflection positivity at scale \(k\) is the assertion that
\[
  c^\ast G_{k,\mathcal F}c\ge0
  \qquad
  \text{for every }c\in\mathbb C^m .
\]
If the window is part of a reconstruction-strength RG estimate, the entries
of \(G_{k,\mathcal F}\) converge to entries
\[
  G_{\ast,\mathcal F;ij}
  =
  S_\ast\bigl((\Theta F_i)F_j\bigr).
\]
For each fixed \(c\), continuity of the finite quadratic form gives
\[
  c^\ast G_{\ast,\mathcal F}c
  =
  \lim_{k\to\infty}c^\ast G_{k,\mathcal F}c
  \ge0 .
\]
Thus positivity passes to the limit for this finite family.  The quantitative
finite-window form is also useful.  If \(G_{k,\mathcal F}\) is Hermitian,
\(G_{k,\mathcal F}\ge \ell\,1_m\), and the limiting-window error satisfies
\[
  |G_{\ast,\mathcal F;ij}-G_{k,\mathcal F;ij}|\le\epsilon
  \qquad
  (1\le i,j\le m),
\]
then
\begin{equation}
  c^\ast G_{\ast,\mathcal F}c
  \ge
  \bigl(\ell-m\epsilon\bigr)\|c\|_2^2 .
  \label{eq:rg-finite-os-positivity-bound}
\end{equation}
Indeed,
\[
  |c^\ast(G_{\ast,\mathcal F}-G_{k,\mathcal F})c|
  \le
  \epsilon\Bigl(\sum_{i=1}^m |c_i|\Bigr)^2
  \le
  m\epsilon\,\|c\|_2^2 .
\]
Equation~\eqref{eq:rg-finite-os-positivity-bound} is a finite
Gram-matrix lower bound, not an OS reconstruction theorem.  For this
entrywise form to give a positive lower bound on a family of size \(m\), one
needs \(\epsilon<\ell/m\).  Hence a directed-family OS theorem cannot be
obtained by fixing an error tolerance independently of the number of test
vectors: if \(\mathcal F_N\) is an increasing family with
\(m_N=|\mathcal F_N|\), the estimate must supply
\(\epsilon_{N,k}\) and \(\ell_{N,k}\) with
\(\ell_{N,k}-m_N\epsilon_{N,k}\ge0\) in the limit being taken, or else use a
stronger operator-norm estimate
\(\|G_{\ast,\mathcal F_N}-G_{k,\mathcal F_N}\|_{\ell^2\to\ell^2}
\le\epsilon^{\rm op}_{N,k}\) with
\(\ell_{N,k}-\epsilon^{\rm op}_{N,k}\ge0\).  Producing such
family-size-compatible error control from RG estimates is a load-bearing
part of the theorem.  OS positivity for the continuum hierarchy requires
this bound, or the corresponding limiting argument, for every finite
positive-time test family in the directed observable germ.  Agreement of
separated two-point entries, critical exponents, or a finite list of
amplitudes does not supply the missing Gram windows.

\begin{lemma}[Directed OS-positive form assembly]
\label{lem:rg-directed-os-positive-form-assembly}
Let
\[
  P_1\subset P_2\subset\cdots
\]
be finite-dimensional subspaces of the positive-time polynomial test algebra,
with dense algebraic union \(P_\infty\) in the positive-time test topology
used for OS reconstruction.  Let \(N_k\to\infty\).  For every
\(N\le N_k\), suppose the regulated Schwinger hierarchy supplies a Hermitian
sesquilinear form
\[
  Q_{k,N}:P_N\times P_N\longrightarrow\mathbb C
\]
such that:
\begin{enumerate}
  \item \(Q_{k,M}|_{P_N\times P_N}=Q_{k,N}\) whenever
  \(N\le M\le N_k\);
  \item \(Q_{k,N}(F,F)\ge0\) for every \(F\in P_N\);
  \item after choosing any norm \(\|\cdot\|_N\) on \(P_N\),
  \begin{equation}
    \eta_{N,k}
    =
    \sup_{\ell\ge k}
    \sup_{\substack{\|F\|_N\le1\\ \|G\|_N\le1}}
    |Q_{k,N}(F,G)-Q_{\ell,N}(F,G)|
    \longrightarrow0
    \label{eq:directed-os-form-cauchy-control}
  \end{equation}
  for each fixed \(N\), with the supremum restricted to scales for which
  \(N\le N_k,N_\ell\).
\end{enumerate}
Then there is a unique compatible positive Hermitian form
\[
  Q_\ast:P_\infty\times P_\infty\longrightarrow\mathbb C
\]
whose restriction to \(P_N\) is the limit of \(Q_{k,N}\) for every fixed
\(N\).  If, in addition, the limiting forms obey the continuity estimate
required by the OS chapter on a dense positive-time family, then \(Q_\ast\)
extends to the positive-time OS pre-Hilbert form.  The lemma by itself does
not supply strong continuity of the time-translation semigroup or the OS-II
growth hypothesis.
\end{lemma}

\begin{proof}
For fixed \(N\), the space of sesquilinear forms on \(P_N\) is finite
dimensional and complete in the operator norm induced by \(\|\cdot\|_N\).
Condition~\eqref{eq:directed-os-form-cauchy-control} therefore gives a
limit form \(Q_{\ast,N}\) on \(P_N\).  Positivity passes to the limit because
\[
  Q_{\ast,N}(F,F)=\lim_{k\to\infty} Q_{k,N}(F,F)\ge0
\]
for each fixed \(F\in P_N\).  Hermiticity passes in the same way.

If \(N\le M\), finite-scale compatibility gives
\[
  Q_{k,M}|_{P_N\times P_N}=Q_{k,N}
\]
for all sufficiently large \(k\).  Passing to the finite-dimensional limits
on \(P_M\) and \(P_N\) gives
\[
  Q_{\ast,M}|_{P_N\times P_N}=Q_{\ast,N}.
\]
Thus one defines \(Q_\ast(F,G)=Q_{\ast,N}(F,G)\) whenever
\(F,G\in P_N\); the preceding compatibility makes this independent of
\(N\).  Positivity and Hermiticity on \(P_\infty\) follow by choosing a
single \(P_N\) that contains the finitely many vectors involved.  The final
extension statement is exactly the continuity step in OS reconstruction:
without a bound in the positive-time topology, a positive algebraic form on
the dense union need not define a continuous form on the desired test
completion.
\end{proof}

The quantitative entrywise estimate
\eqref{eq:rg-finite-os-positivity-bound} is one way to prove the Cauchy and
positivity inputs of Lemma~\ref{lem:rg-directed-os-positive-form-assembly}.
For a nested family with \(m_N\) basis vectors it must be used with a
scale-dependent error schedule satisfying
\[
  \liminf_{k\to\infty}\bigl(\ell_{N,k}-m_N\epsilon_{N,k}\bigr)\ge0
  \qquad\text{for each fixed }N.
\]
If the goal is a uniform lower bound on an increasing family, the right
condition is correspondingly uniform in \(N\); a fixed entrywise tolerance is
not enough.  This is the precise mathematical content behind the warning
that finite OS Gram windows must be controlled as a directed family.

\paragraph{When universality reconstructs the same local QFT.}
Suppose a Wilsonian universality statement has a local-QFT-strength
observable germ and that the normalized reconstructions of all microscopic
regulators converge to a common germ \(O_\ast\).  The common germ then
determines a common Schwinger or Wightman hierarchy in the sense of the
chosen reconstruction framework.  Applying the reconstruction theorem
proved earlier in the monograph gives the corresponding Hilbert space,
vacuum vector, fields or observable algebra, and spacetime symmetry
representation.  Thus the phrase ``the regulators define the same local QFT''
means: after the declared normalizations, their full local-QFT-strength
observable germs converge to one reconstruction input, and the reconstructed
presentations are identified by the uniqueness clause of that reconstruction
theorem.  Contact-term choices are part of the source-local extension data;
they may change extensions across collision diagonals without changing the
separated-point germ.

This paragraph is not an additional fixed-point theorem.  It records the
dependency chain that a model-specific universality theorem must establish:
RG attraction in a Banach chart, convergence of the full projective
observable germ, verification of OS or Wightman positivity and growth
hypotheses, and then reconstruction.  A calculation of critical exponents,
a finite bootstrap table, a finite tensor-window comparison, or a lattice
finite-size scaling plot supplies only the finite windows it actually
controls.

The finite-window form is still useful because it is the object produced by
simulations, rigorous finite-volume estimates, and perturbative
calculations.  If \(\mathfrak A_0\subset\mathfrak A\) is finite and, for
each \(\alpha\in\mathfrak A_0\),
\begin{equation}
  p_\alpha\!\left(
    g_i\,\operatorname{Rec}_{i,n}^{\alpha}(V_i)
    -
    (O_\ast)_\alpha
  \right)
  \leq e_{\alpha,i,n},
  \qquad
  p_\alpha\!\left(
    g_j\,\operatorname{Rec}_{j,n}^{\alpha}(V_j)
    -
    (O_\ast)_\alpha
  \right)
  \leq e_{\alpha,j,n},
\end{equation}
then the triangle inequality gives the finite-window estimate
\begin{equation}
  p_\alpha\!\left(
    g_i\,\operatorname{Rec}_{i,n}^{\alpha}(V_i)
    -
    g_j\,\operatorname{Rec}_{j,n}^{\alpha}(V_j)
  \right)
  \leq e_{\alpha,i,n}+e_{\alpha,j,n},
  \qquad \alpha\in\mathfrak A_0 .
  \label{eq:finite-window-universality-estimate}
\end{equation}

A finite-window estimate becomes substantially more informative when its
error is expressed in RG coordinates.  Fix a finite window
\(\alpha\in\mathfrak A\) and a scale \(n_0\) after which two normalized
microscopic data lie in a common stable chart at \(V_\ast\).  Write
\(u\in E_u\) for the relevant coordinate and \(s\in E_s\) for the stable
coordinate in that chart.  Suppose that, after the allowed normalizations,
the relevant mismatch is bounded by
\[
  \|u_i(n_0)-u_j(n_0)\|\leq\delta_u ,
\]
and that the stable mismatch \(d_\ell\) after \(\ell\) further RG steps obeys
the finite recurrence
\begin{equation}
  d_{\ell+1}
  \leq
  \theta d_\ell+\eta_\ell,
  \qquad
  d_0\leq\delta_s,\qquad 0\leq\theta<1 .
  \label{eq:finite-rg-window-stable-recurrence}
\end{equation}
Here \(\eta_\ell\) is the one-step comparison defect: it may come from
different microscopic blocking kernels, a finite auxiliary-to-target
intertwining error, or a controlled truncation of the stable coordinate.
Assume also that the reconstructed \(\alpha\)-window is Lipschitz in the
chart, with constants \(C_{u,\alpha}\) and \(C_{s,\alpha}\), and that the
source-tail and normalization errors in this window at depth \(m\) are
bounded by \(\Xi_{\alpha,m}\).  Then the finite observable-window bound is
\begin{equation}
\begin{aligned}
  p_\alpha\!\left(
    g_i\,\operatorname{Rec}_{i,n_0+m}^{\alpha}(V_i)
    -
    g_j\,\operatorname{Rec}_{j,n_0+m}^{\alpha}(V_j)
  \right)
  &\leq
  C_{u,\alpha}\delta_u
  +
  C_{s,\alpha}\theta^m\delta_s                                      \\
  &\quad+
  C_{s,\alpha}\sum_{\ell=0}^{m-1}
    \theta^{m-1-\ell}\eta_\ell
  +
  \Xi_{\alpha,m}.
\end{aligned}
  \label{eq:stable-chart-finite-window-bound}
\end{equation}
This is the finite-scale form of a Wilsonian universality comparison.  Its
derivation is the iteration of
\eqref{eq:finite-rg-window-stable-recurrence}, followed by the Lipschitz
bound for the reconstruction map in the chosen window.  Exact tuning means
\(\delta_u=0\).  A theorem-level universality result must prove constants
for every window in a directed family and show that the right-hand side of
\eqref{eq:stable-chart-finite-window-bound} tends to zero along the
claimed continuum or long-distance limit.  Without that directed-window
limit, \eqref{eq:stable-chart-finite-window-bound} remains a precise
finite observation bound.

The theorem-level burden is the passage from these finite bounds to
\eqref{eq:observable-germ-universality-convergence} for the directed family
of observation windows.

\begin{remark}[Universality as an equivalence relation after reconstruction]
\label{rem:universality-equivalence-relation}
Assume the data of Definition~\ref{def:wilsonian-universality-statement}, and
suppose the displayed convergence to a common \(O_\ast\) holds.  Define
\(i\sim j\) if there exist \(g_i,g_j\in G\) such that
\[
  g_i\,\operatorname{Rec}_{i,n}(V_i)
  -
  g_j\,\operatorname{Rec}_{j,n}(V_j)
  \longrightarrow0
  \qquad\text{in }\mathcal O .
\]
Then \(\sim\) is an equivalence relation on \(I\).
Reflexivity holds by taking the same normalization on both sides.  Symmetry
is immediate by changing the sign of the displayed difference in the vector
space \(\mathcal O\).

For transitivity, suppose \(i\sim j\) and \(j\sim k\).  Then there are
normalizations \(g_i,g_j,h_j,h_k\in G\) for which
\[
  g_i\operatorname{Rec}_{i,n}(V_i)
  -
  g_j\operatorname{Rec}_{j,n}(V_j)\to0,
  \qquad
  h_j\operatorname{Rec}_{j,n}(V_j)
  -
  h_k\operatorname{Rec}_{k,n}(V_k)\to0 .
\]
In the theorem-level situation above, each normalized sequence also converges
to the same orbit representative \(O_\ast\) after some allowed normalization.
Choose the normalizations \(a_i,a_j,a_k\in G\) that appear in
Definition~\ref{def:wilsonian-universality-statement}.  Then
\[
  a_i\operatorname{Rec}_{i,n}(V_i)\to O_\ast,
  \qquad
  a_k\operatorname{Rec}_{k,n}(V_k)\to O_\ast,
\]
so
\[
  a_i\operatorname{Rec}_{i,n}(V_i)
  -
  a_k\operatorname{Rec}_{k,n}(V_k)\to0 .
\]
Thus \(i\sim k\).  The proof uses the existence of a common reconstructed
limit and therefore controls more than pairwise similarity of formal coupling
lists.
\end{remark}

Thus a universality class may be summarized by a triple
\[
  (\mathcal R,\mathcal N,V_\ast)
\]
together with the observable topology, reconstruction maps, and normalization
group of Definition~\ref{def:wilsonian-universality-statement}.  Symmetry and
dimension may guide the choice of \(I\) and the list of relevant coordinates,
but they do not by themselves define the class.  The theorem is the attraction
and reconstruction statement.

\paragraph{Correction-to-scaling data.}
The limiting reconstructed datum is only the first layer of a Wilsonian
universality theorem.  A sharper theorem also identifies the controlled
approach to the limit.  This matters because the amplitudes of irrelevant
directions depend on the microscopic regulator, while the decay exponents and
the reconstructed correction distributions are fixed by the RG chart and the
fixed point.

\begin{definition}[Correction-to-scaling datum]
\label{def:correction-to-scaling-datum}
Assume the data of
Definition~\ref{def:wilsonian-universality-statement}.  A
correction-to-scaling datum of order \(A\) consists of:
\begin{enumerate}
\item a locally convex topology on \(\mathcal O\), specified by seminorms
\(\{p_\beta\}\);
\item a stable coordinate chart near \(V_\ast\) in which the tuned
trajectories have stable coordinates \(w_i(n)\in E_s\);
\item vectors \(e_1,\ldots,e_A\in E_s\) and positive numbers
\(\omega_1\le\cdots\le\omega_A\) such that the linearized RG map sends
\[
  e_a\longmapsto L^{-\omega_a}e_a ;
\]
\item a number \(\omega_{A+1}>\omega_A\), or another declared remainder
rate, bounding the omitted stable directions;
\item constants \(c_{i,a}\in\mathbb C\), depending on the microscopic
regulator \(i\), and remainders \(\rho_i(n)\in E_s\) satisfying
\begin{equation}
  w_i(n)
  =
  \sum_{a=1}^{A}c_{i,a}L^{-n\omega_a}e_a+\rho_i(n),
  \qquad
  \|\rho_i(n)\|_{E_s}
  \le
  C_i\bigl(L^{-n\omega_{A+1}}+L^{-2n\omega_1}\bigr);
  \label{eq:correction-stable-coordinate-expansion}
\end{equation}
\item differentiability of the normalized reconstruction map at \(V_\ast\),
with first derivative
\[
  D\operatorname{Rec}_{\ast}:E_s\longrightarrow\mathcal O
\]
and a quadratic remainder estimate in every seminorm \(p_\beta\).
\end{enumerate}
The associated correction distributions are
\[
  \mathcal C_a=D\operatorname{Rec}_{\ast}(e_a)\in\mathcal O .
\]
\end{definition}

Under these hypotheses the normalized reconstructed observables have the
expansion
\begin{equation}
  g_i\operatorname{Rec}_{i,n}(V_i)
  =
  O_\ast+
  \sum_{a=1}^{A}c_{i,a}L^{-n\omega_a}\mathcal C_a
  +
  \mathcal E_i(n),
  \label{eq:correction-to-scaling-observable-expansion}
\end{equation}
where, for every seminorm \(p_\beta\), there is a constant
\(C_{\beta,i}\) such that
\begin{equation}
  p_\beta(\mathcal E_i(n))
  \le
  C_{\beta,i}\bigl(L^{-n\omega_{A+1}}+L^{-2n\omega_1}\bigr).
  \label{eq:correction-to-scaling-observable-remainder}
\end{equation}
Indeed, write the normalized reconstruction map in the stable chart as
\[
  \operatorname{Rec}_{\ast}(V_\ast+w)
  =
  O_\ast+D\operatorname{Rec}_{\ast}(w)+Q(w),
  \qquad
  p_\beta(Q(w))\le C_\beta\|w\|_{E_s}^2
\]
for \(w\) in a sufficiently small stable neighborhood.  Substitution of
\eqref{eq:correction-stable-coordinate-expansion} gives the linear
correction terms and puts the stable-coordinate remainder together with
\(Q(w_i(n))\) into
\eqref{eq:correction-to-scaling-observable-remainder}.

This is the precise content behind the common phrase ``same universality
class but different corrections to scaling.''  It is not a theorem about a
model until the stable-coordinate expansion, the reconstruction derivative,
and the seminorm estimates have all been proved in the chosen RG framework.
The coefficients \(c_{i,a}\) are microscopic coordinates.  The exponents
\(\omega_a\) and correction distributions \(\mathcal C_a\) are fixed-point
data in the declared Wilsonian chart and observable topology.

\begin{openproblem}[Nonperturbative Wilsonian fixed points]
\label{op:constructive-rg-nonperturbative-fixed-points}
Construct non-Gaussian Wilsonian fixed points in three and four dimensions as
fixed points of explicit RG maps on Banach spaces of interactions, with
stable manifolds, operator spectra, and reconstruction of the associated
local QFT data.  The open directions include: short-range scalar critical
fixed points with distributional reconstruction; gauge-compatible fixed
points in the lattice or BV sense of
Definition~\ref{def:gauge-compatible-wilsonian-rg-structure}; and estimates that
identify the limiting local, extended, and cohomological observables rather
than only the local coordinate flow.
\end{openproblem}
