\chapter{Wilson Lattice Gauge Theory}
\label{ch:constructive-wilson-lattice-gauge-theory}

\ConventionsEuclidean

Wilson lattice gauge theory gives a gauge-invariant finite-regulator
definition of compact gauge fields.  Its continuum use requires a scaling
trajectory, reflection positivity, locality, and a reconstruction or
continuum-limit argument.  The lattice theory is not merely a discretized
notation for continuum perturbation theory; it is a nonperturbative
regulated path integral.

\section{Link Variables and Gauge Transformations}

Let \(\Lambda\subset a\mathbb Z^D\) be a finite hypercubic lattice and \(G\)
a compact Lie group.  To each oriented link \(\ell=(x,x+a\hat\mu)\), assign
\[
  U_\mu(x)\in G,
\]
with \(U_{-\mu}(x+a\hat\mu)=U_\mu(x)^{-1}\).  A lattice gauge transformation
is a function \(g:\Lambda^0\to G\) acting by
\[
  U_\mu(x)\mapsto g(x)U_\mu(x)g(x+a\hat\mu)^{-1}.
\]
The plaquette holonomy is
\[
  U_{\mu\nu}(x)
  =
  U_\mu(x)U_\nu(x+a\hat\mu)
  U_\mu(x+a\hat\nu)^{-1}U_\nu(x)^{-1}.
\]

\section{Wilson Action}

Fix a faithful unitary representation \(R\) and normalize the trace in the
continuum convention of the gauge-theory volume.  A Wilson plaquette action
has the form
\[
  S_W[U]
  =
  \beta\sum_{p}
  \left(
  1-\frac{1}{\dim R}\operatorname{Re}\operatorname{Tr}_R U_p
  \right).
\]
The finite-volume partition function is the compact group integral
\[
  Z=\int\prod_{\ell}dU_\ell\,e^{-S_W[U]},
\]
where \(dU\) is Haar measure.

\section{A Finite \texorpdfstring{\(\mathbb Z_2\)}{Z2} Gauge Benchmark}

Before taking any continuum limit, a lattice gauge theory is an ordinary
finite-dimensional integral.  The smallest nontrivial pure gauge benchmark is
the \(\mathbb Z_2\) gauge model on a finite periodic cubic lattice.  Let
\(E\) be the set of oriented positive links, let \(P\) be the set of
elementary plaquettes, and assign
\[
  U_e\in\{\pm1\},\qquad e\in E .
\]
For a plaquette \(p\), define
\[
  U_p=\prod_{e\in\partial p}U_e ,
\]
where orientation is irrelevant because \(U_e^{-1}=U_e\).  The finite
partition function at parameter \(\beta\in\mathbb R\) is
\[
  Z_\Lambda(\beta)
  =
  \sum_{U:E\to\{\pm1\}}
  \exp\!\left(\beta\sum_{p\in P}U_p\right).
\]
A gauge transformation is a function \(g:V\to\{\pm1\}\) acting on a link
\(e=(x,y)\) by
\[
  U_e\longmapsto g_xU_eg_y .
\]
The plaquette variables and Wilson loop variables
\[
  W(C)=\prod_{e\in C}U_e
\]
for closed lattice contours \(C\) are gauge invariant.

\begin{proposition}[Finite \(\mathbb Z_2\) character expansion]
\label{prop:z2-gauge-character-expansion}
Let \(t=\tanh\beta\).  For a closed contour \(C\), regard \(C\) as a
\(\mathbb Z_2\)-one-chain, and regard a subset \(A\subset P\) as a
\(\mathbb Z_2\)-two-chain with boundary \(\partial A\).  Then
\[
  Z_\Lambda(\beta)
  =
  2^{|E|}(\cosh\beta)^{|P|}
  \sum_{\substack{A\subset P\\ \partial A=0}} t^{|A|},
\]
and
\[
  \langle W(C)\rangle_{\Lambda,\beta}
  =
  \frac{\displaystyle
  \sum_{\substack{A\subset P\\ \partial A=C}}t^{|A|}}
  {\displaystyle
  \sum_{\substack{A\subset P\\ \partial A=0}}t^{|A|}} .
\]
The sums are finite sums over plaquette subsets.  Thus the strong-coupling
area mechanism in this model is not an analogy: at finite cutoff the Wilson
loop numerator is exactly a weighted sum over lattice surfaces whose
boundary is the loop.
\end{proposition}

\begin{proof}
For each plaquette variable \(U_p=\pm1\),
\[
  e^{\beta U_p}=\cosh\beta\,(1+tU_p)
  =
  \cosh\beta\sum_{n_p=0}^1 t^{n_p}U_p^{n_p}.
\]
Multiplying over plaquettes gives
\[
  \exp\!\left(\beta\sum_p U_p\right)
  =
  (\cosh\beta)^{|P|}
  \sum_{A\subset P}
  t^{|A|}
  \prod_{p\in A}\prod_{e\in\partial p}U_e .
\]
Summing over a single link variable gives
\(\sum_{U_e=\pm1}U_e^m=0\) if \(m\) is odd and \(2\) if \(m\) is even.
Therefore the partition function keeps precisely those plaquette subsets for
which every link appears an even number of times, namely
\(\partial A=0\), and contributes the overall factor \(2^{|E|}\).

For the Wilson-loop numerator, the additional factor
\(\prod_{e\in C}U_e\) changes the parity condition on each link from
\((\partial A)_e=0\) to \((\partial A)_e=C_e\).  Dividing the resulting
finite numerator by the finite partition function gives the displayed
expectation value.
\end{proof}

The companion script
\[
  \texttt{qft\_scripts/z2\_gauge\_3d\_metropolis.py}
\]
samples this finite model by single-link Metropolis updates and measures the
plaquette average and rectangular Wilson loops.  Its output is finite-cutoff
data.  An infinite-volume transition, a continuum limit, or a continuum
confinement statement requires additional limiting hypotheses and estimates.

\section{Surface Counting and the Strong-Coupling Area Bound}

The exact formula in
Proposition~\ref{prop:z2-gauge-character-expansion} becomes an area-law
estimate only after one controls how many surfaces of a given area have the
same boundary.  This counting step is often hidden in the phrase
``strong-coupling expansion.''  We spell out the finite statement.

Fix a finite lattice and a closed contour \(C\) which is a boundary in the
\(\mathbb Z_2\) plaquette chain complex.  Define the lattice area
\[
  A_{\min}(C)
  =
  \min\{|A|:A\subset P,\ \partial A=C\}.
  \label{eq:z2-minimal-lattice-area}
\]
For \(n\geq0\), let
\[
  N_C(n)
  =
  \#\{A\subset P:\partial A=C,\ |A|=A_{\min}(C)+n\}.
  \label{eq:z2-surface-counting-number}
\]
Similarly let \(N_0(n)\) count closed plaquette subsets of area \(n\).

\begin{proposition}[Finite surface expansion and area estimate]
\label{prop:z2-surface-counting-area-bound}
Let \(\beta\geq0\) and \(t=\tanh\beta\).  If \(C\) is a boundary, then
\[
  \langle W(C)\rangle_{\Lambda,\beta}
  =
  t^{A_{\min}(C)}
  \frac{\sum_{n\geq0}N_C(n)t^n}
       {\sum_{n\geq0}N_0(n)t^n}.
  \label{eq:z2-surface-counting-expansion}
\]
In particular, since the denominator is at least \(1\),
\[
  0\leq
  \langle W(C)\rangle_{\Lambda,\beta}
  \leq
  t^{A_{\min}(C)}
  \sum_{n\geq0}N_C(n)t^n .
  \label{eq:z2-area-bound-before-entropy}
\]
If, for a family of contours, the surface entropy obeys
\[
  N_C(n)\leq K(C)\rho^n
  \qquad(n\geq0)
  \label{eq:z2-surface-entropy-hypothesis}
\]
with \(0\leq\rho t<1\), then
\[
  \langle W(C)\rangle_{\Lambda,\beta}
  \leq
  \frac{K(C)}{1-\rho t}\,
  \exp\{-[-\log t]A_{\min}(C)\}.
  \label{eq:z2-entropy-area-bound}
\]
\end{proposition}

\begin{proof}
The numerator in
Proposition~\ref{prop:z2-gauge-character-expansion} is a finite sum over
plaquette subsets satisfying \(\partial A=C\).  Grouping these subsets by
the excess area \(n=|A|-A_{\min}(C)\) gives the numerator
\[
  t^{A_{\min}(C)}\sum_{n\geq0}N_C(n)t^n .
  \]
The same grouping with \(C=0\) gives the denominator.  For
\(\beta\geq0\), all terms in the denominator are nonnegative and the empty
closed surface gives the term \(1\); hence the denominator is at least
\(1\), proving \eqref{eq:z2-area-bound-before-entropy}.  The final estimate
is the geometric-series bound
\[
  \sum_{n\geq0}N_C(n)t^n
  \leq
  K(C)\sum_{n\geq0}(\rho t)^n
  =
  \frac{K(C)}{1-\rho t}.
  \]
\end{proof}

For a rectangular loop \(C_{R,T}\) in a coordinate plane,
\(A_{\min}(C_{R,T})=RT/a^2\) when the rectangle is smaller than the chosen
finite box and the flat spanning rectangle is the unique minimal surface in
that homology class.  Equation~\eqref{eq:z2-entropy-area-bound} then gives a
finite-regulator area bound at sufficiently small \(t\).  If the constants
in the surface-counting estimate have at most perimeter growth,
\(K(C_{R,T})\leq \exp\{c(R+T)/a\}\), the area coefficient in lattice units is
bounded below by
\[
  \sigma_{\rm lat}(\beta)\geq -\log(\tanh\beta)
\]
with the displayed perimeter and finite entropy prefactors kept separately.
This is a strong-coupling theorem at fixed cutoff.  It is not a proof of
continuum confinement, because a continuum Yang--Mills limit in four
dimensions would require control along a scaling trajectory where the Wilson
parameter does not remain in this small-\(\beta\) domain.

\begin{example}[One-cube surface polynomial]
\label{ex:z2-one-cube-surface-polynomial}
On the boundary of one elementary cube, let \(C\) be the boundary of one
face.  There are exactly two plaquette subsets with boundary \(C\): the face
itself, with area \(1\), and the complementary five faces, with area \(5\).
The closed subsets are the empty surface and the six-face cube boundary.
Therefore
\[
  \langle W(C)\rangle
  =
  \frac{t+t^5}{1+t^6}.
  \label{eq:z2-one-cube-wilson-polynomial}
\]
This example shows explicitly why the leading strong-coupling term is the
minimal spanning surface and why closed surfaces correct both numerator and
denominator.
\end{example}

\section{Compact-Group Character Expansion}
\label{sec:compact-group-character-expansion}

The \(\mathbb Z_2\) expansion has a nonabelian analogue.  The analogue is
not a sum over unoriented plaquette subsets; it is a finite tensor-network
sum over irreducible representations and Haar projectors.  This distinction
is important: in a nonabelian gauge theory the local degrees of freedom
carried by a surface include representation labels and invariant tensors at
links.

Let \(G\) be a compact Lie group and let \(\widehat G\) be the set of
equivalence classes of finite-dimensional irreducible unitary
representations.  For \(\lambda\in\widehat G\), write
\[
  D_\lambda:G\to U(V_\lambda),
  \qquad
  \chi_\lambda(U)=\operatorname{Tr}_{V_\lambda}D_\lambda(U).
\]
Let \(K:G\to\mathbb C\) be a continuous class function.  Its character
coefficients are
\[
  c_\lambda[K]
  =
  \int_G dU\,K(U)\overline{\chi_\lambda(U)} .
  \label{eq:compact-character-coefficients}
\]
Peter--Weyl orthogonality gives the \(L^2(G)\) expansion
\[
  K(U)=\sum_{\lambda\in\widehat G}c_\lambda[K]\chi_\lambda(U),
  \label{eq:compact-character-expansion}
\]
with rapid coefficient decay when \(K\) is smooth.  For the Wilson plaquette
weight with the irrelevant constant removed one may take
\[
  K_\beta(U)=
  \exp\!\left(
    \frac{\beta}{\dim R}\operatorname{Re}\operatorname{Tr}_R U
  \right).
  \label{eq:compact-wilson-plaquette-weight}
\]

\begin{proposition}[Compact-group spin-foam expansion at finite cutoff]
\label{prop:compact-group-spin-foam-expansion}
On a finite lattice, replace each plaquette Boltzmann factor by a smooth
class function \(K(U_p)\).  Then the finite-volume partition function is
represented by
\[
  Z_\Lambda[K]
  =
  \sum_{\lambda:P\to\widehat G}
  \left(\prod_{p\in P}c_{\lambda_p}[K]\right)
  \mathcal A_\Lambda(\{\lambda_p\}),
  \label{eq:compact-spin-foam-partition}
\]
where \(\mathcal A_\Lambda(\{\lambda_p\})\) is the tensor contraction
obtained by assigning \(V_{\lambda_p}\) or \(V_{\lambda_p}^*\) to the four
oriented sides of each plaquette and, for every link \(\ell\), inserting
the Haar projector
\[
  P_\ell(\{\lambda_p\}_{p\supset\ell})
  =
  \int_G dU_\ell
  \bigotimes_{p\supset\ell}
  D_{\lambda_p}(U_\ell)^{\epsilon(p,\ell)} ,
  \label{eq:compact-link-haar-projector}
\]
with \(\epsilon(p,\ell)=+1\) if the plaquette orientation uses \(U_\ell\)
and \(\epsilon(p,\ell)=-1\) if it uses \(U_\ell^{-1}\).  A Wilson loop in
representation \(S\) is represented by inserting the additional boundary
tensor \(D_S(U_\ell)^{\pm1}\) on each link of the loop before taking the
same Haar projectors.
\end{proposition}

\begin{proof}
Insert the character expansion
\eqref{eq:compact-character-expansion} independently for every plaquette.
After choosing bases of the representation spaces, every plaquette character
is a cyclic contraction of four matrix factors, one for each link in its
boundary.  Each link variable appears only in the factors associated to
plaquettes incident on that link, and in the Wilson-loop insertion if the
link belongs to the loop.  Integrating over this link gives precisely the
Haar average
\eqref{eq:compact-link-haar-projector}.  Performing the finitely many link
integrals leaves a finite tensor contraction for each plaquette-label
assignment.  Truncating the representation labels gives finite tensor
network sums.  For smooth \(K\), in particular for
\eqref{eq:compact-wilson-plaquette-weight}, the character coefficients
decay rapidly enough that the truncated sums converge to the original
finite-dimensional integral.  The \(L^2\) Peter--Weyl statement gives the
same formula for less regular class functions after the corresponding
convergence mode has been specified.
\end{proof}

For \(G=SU(N)\), \(N\ge3\), with \(R\) the fundamental representation and
with the conventional Wilson weight
\[
  K_\beta(U)=
  \exp\!\left(\frac{\beta}{N}\operatorname{Re}\operatorname{Tr}U\right),
  \label{eq:suN-fundamental-wilson-weight}
\]
the small-\(\beta\) coefficients of the fundamental and antifundamental
characters are
\[
  c_F[K_\beta]=c_{\overline F}[K_\beta]
  =
  \frac{\beta}{2N}+O(\beta^2).
  \label{eq:suN-fundamental-character-coefficient}
\]
This follows by expanding \(K_\beta=1+\frac{\beta}{2N}(\chi_F+\chi_{\bar F})
 +O(\beta^2)\) and using character orthogonality.  The coefficient changes
for \(SU(2)\), because the fundamental representation is self-conjugate and
\(\operatorname{Re}\chi_F=\chi_F\).

\begin{example}[Fundamental plaquette normalization in \(SU(N)\)]
\label{ex:suN-fundamental-plaquette-strong-coupling-normalization}
For \(N\ge3\), let
\[
  w_p(U)=\frac1N\operatorname{Re}\operatorname{Tr}U_p .
\]
At \(\beta=0\), the plaquette holonomy \(U_p\) is Haar distributed, so
\(\langle w_p\rangle_{\beta=0}=0\).  Differentiating the normalized
expectation at \(\beta=0\) gives
\[
  \left.\frac{\dd}{\dd\beta}\langle w_p\rangle_\beta\right|_{\beta=0}
  =
  \int_G dU\,
  \frac{\chi_F(U)+\chi_{\bar F}(U)}{2N}\,
  \frac{\chi_F(U)+\chi_{\bar F}(U)}{2N}
  =
  \frac{1}{2N^2}.
  \label{eq:suN-one-plaquette-slope}
\]
Thus, for \(SU(3)\),
\[
  \langle w_p\rangle_\beta
  =
  \frac{\beta}{18}+O(\beta^2).
  \label{eq:su3-one-plaquette-slope}
\]
The calculation fixes the normalization used in the nonabelian
strong-coupling expansion.  Higher-area Wilson loops are obtained by the
same Haar-projector contractions, but a theorem-level area estimate also
requires bounds on representation sums, invariant-tensor contractions, and
surface entropy.
\end{example}

\section{Continuum Expansion}

For a smooth continuum connection \(A_\mu(x)\), set
\[
  U_\mu(x)=\exp(a A_\mu(x)).
\]
The plaquette has expansion
\[
  U_{\mu\nu}(x)
  =
  \exp\!\left(a^2F_{\mu\nu}(x)+O(a^3)\right).
\]
If the representation trace obeys
\[
  \operatorname{Tr}_R(XY)=\kappa_R\,\operatorname{tr}(XY),
\]
then
\[
  1-\frac{1}{\dim R}\operatorname{Re}\operatorname{Tr}_R U_{\mu\nu}
  =
  -\frac{a^4\kappa_R}{2\dim R}
  \operatorname{tr}(F_{\mu\nu}F_{\mu\nu})
  +O(a^6)
\]
for anti-Hermitian \(A_\mu\).  Matching to
\[
  \frac{1}{4g^2}\int d^Dx\,\operatorname{tr}(F_{\mu\nu}F_{\mu\nu})
\]
fixes the relation between \(\beta\), \(g\), and the trace normalization.

\section{Improved Gauge Actions at Tree Level}
\label{sec:lattice-tree-level-improved-gauge-actions}

The lattice action is part of the regulator definition.  Two local
gauge-invariant actions may define the same continuum theory only after a
scaling trajectory and a comparison of renormalized coordinates have been
specified.  Improvement is the finite-regulator problem of choosing the
local action so that selected irrelevant coordinates are absent, or smaller,
in a chosen coordinate system.

For \(G=SU(N)\) in the fundamental trace convention used above, let
\(U_{\mu\nu}^{1\times1}(x)\) be the elementary plaquette and let
\(U_{\mu\nu}^{1\times2}(x)\) denote the oriented rectangle with side lengths
\(a\) and \(2a\) in the \(\mu,\nu\) plane.  The plaquette-plus-rectangle
family is
\[
  S_{c_0,c_1}[U]
  =
  \frac{\beta}{N}
  \sum_x\sum_{\mu<\nu}
  \left[
    c_0\bigl(N-\operatorname{Re}\operatorname{Tr}U_{\mu\nu}^{1\times1}(x)\bigr)
    +c_1\sum_{\rho\in\{(\mu,\nu),(\nu,\mu)\}}
     \bigl(N-\operatorname{Re}\operatorname{Tr}U_\rho^{1\times2}(x)\bigr)
  \right].
  \label{eq:plaquette-rectangle-gauge-action}
\]
The two rectangles in the last sum have their long side in the first or
second direction of the ordered pair.  Since a \(1\times2\) rectangle has
twice the area of a plaquette, its leading \(F_{\mu\nu}^2\) contribution is
four times the plaquette contribution.  Thus the continuum normalization is
\[
  c_0+8c_1=1.
  \label{eq:plaquette-rectangle-continuum-normalization}
\]
The Wilson action is \(c_0=1,c_1=0\).

\begin{proposition}[Tree-level rectangle improvement]
\label{prop:tree-level-rectangle-gauge-improvement}
In the plaquette-plus-rectangle family
\eqref{eq:plaquette-rectangle-gauge-action}, the tree-level cancellation of
the abelian quadratic \(O(a^2)\) derivative artifact is equivalent to
\[
  c_0+20c_1=0.
  \label{eq:plaquette-rectangle-derivative-cancellation}
\]
Together with \eqref{eq:plaquette-rectangle-continuum-normalization}, this
gives
\[
  c_1=-\frac{1}{12},\qquad c_0=\frac{5}{3}.
  \label{eq:tree-level-symanzik-gauge-coefficients}
\]
\end{proposition}

\begin{proof}
It suffices for the tree-level quadratic calculation to work in an abelian
Cartan direction.  Let \(F=F_{\mu\nu}\) be a smooth scalar field strength and
consider a rectangle centered at \(x\) with side lengths \(ra\) in the
\(\mu\)-direction and \(sa\) in the \(\nu\)-direction.  Its flux is
\[
  \Phi_{r,s}(x)
  =
  \int_{-ra/2}^{ra/2}\!\dd u
  \int_{-sa/2}^{sa/2}\!\dd v\,
  F(x+u\hat\mu+v\hat\nu).
\]
Taylor expansion around the center gives
\[
  \Phi_{r,s}
  =
  r s a^2 F
  +
  \frac{r s a^4}{24}
  \left(r^2\partial_\mu^2+s^2\partial_\nu^2\right)F
  +O(a^6).
  \label{eq:rectangle-flux-expansion}
\]
The small-field loop weight is proportional to \(\frac12\Phi_{r,s}^2\).
After summing over lattice sites, the \(F\partial^2F\) term may be integrated
by parts up to boundary terms or finite-volume periodic identities:
\[
  \frac12\Phi_{r,s}^2
  =
  \frac12r^2s^2a^4F^2
  -
  \frac{r^2s^2a^6}{24}
  \left[
    r^2(\partial_\mu F)^2+s^2(\partial_\nu F)^2
  \right]
  +O(a^8).
  \label{eq:rectangle-flux-square-expansion}
\]
For the plaquette \((r,s)=(1,1)\), the derivative coefficient in one
\(\mu\nu\)-plane is proportional to \(1\).  For the two \(1\times2\)
rectangles, \((r,s)=(1,2)\) and \((2,1)\), the corresponding coefficients
add to
\[
  4(1^2+2^2)=20
\]
in the symmetric combination
\((\partial_\mu F_{\mu\nu})^2+(\partial_\nu F_{\mu\nu})^2\).  Therefore the
quadratic \(O(a^2)\) derivative artifact is proportional to \(c_0+20c_1\).
Setting this coefficient to zero and imposing the leading continuum
normalization \(c_0+8c_1=1\) gives
\eqref{eq:tree-level-symanzik-gauge-coefficients}.
\end{proof}

The preceding proposition is deliberately a tree-level statement.  In a
nonabelian interacting theory, the dimension-six gauge-invariant operator
basis contains covariant-derivative and cubic-curvature terms, modulo Bianchi
identities and field redefinitions.  Loop-level improvement is therefore a
renormalized-coordinate problem whose conditions extend the two tree-level
algebraic equations above.  The full Luescher--Weisz class also allows chair and
parallelogram six-link loops; with coefficients
\((c_0,c_1,c_2,c_3)\) in the standard orientation-counting convention, the
leading normalization is
\[
  c_0+8c_1+16c_2+8c_3=1,
  \label{eq:luescher-weisz-normalization}
\]
where the numerical factors count orientations and leading squared areas in
one coordinate plane.  The commonly used Iwasaki one-parameter action belongs
to the plaquette-plus-rectangle subfamily with \(c_1\simeq -0.331\) and
\(c_0=1-8c_1\); this coefficient is fixed by a renormalization-group
criterion, not by the tree-level cancellation
\eqref{eq:plaquette-rectangle-derivative-cancellation}.

Reflection positivity is an additional constraint.  The compact Euclidean
integral \eqref{eq:plaquette-rectangle-gauge-action} is well-defined whenever
the Boltzmann weight is integrable, but a transfer-matrix reconstruction
requires a positivity argument for the chosen action.  Negative rectangle
coefficients, useful for improvement, need not preserve the elementary
reflection-positivity proof available for the Wilson plaquette action.
Thus improvement, continuum normalization, and Hilbert-space reconstruction
are separate requirements of the regulator datum.

\section{Gauge-Covariant Link Smearing}
\label{sec:gauge-covariant-link-smearing}

Smearing is a regulator-level map from link fields to link fields.  It is
used to build extended observables, improve overlap with low-energy states,
and, in some actions, define the local lattice Lagrangian itself.  Its
mathematical datum is a map
\[
  \mathcal S:\mathcal U_\Lambda=G^E\to G^E
\]
with a locality radius and a gauge-covariance law.  For a link
\(\ell=(x,y)\), gauge covariance means
\[
  \mathcal S_\ell(g\cdot U)
  =
  g(x)\mathcal S_\ell(U)g(y)^{-1}.
  \label{eq:smearing-gauge-covariance}
\]
When a Wilson loop is built from \(\mathcal S(U)\), it is a different
regularized line operator from the loop built from \(U\).  When an action is
built from \(\mathcal S(U)\), it is a different regulator action.  The
continuum comparison is again a scaling-trajectory and operator-scheme
question.

Let \(G=SU(N)\).  The oriented staple sum \(C_\ell(U)\) associated with a
link \(\ell=(x,y)\) is the sum over the \(2(D-1)\) shortest three-link paths
from \(x\) to \(y\) that, together with \(\ell^{-1}\), form an elementary
plaquette.  It transforms as a link:
\[
  C_\ell(g\cdot U)=g(x)C_\ell(U)g(y)^{-1}.
  \label{eq:staple-link-covariance}
\]
The APE preprojection matrix is
\[
  X_\ell(U)
  =
  (1-\alpha)U_\ell
  +
  \frac{\alpha}{2(D-1)}C_\ell(U),
  \qquad \alpha\in\mathbb R .
  \label{eq:ape-preprojection}
\]
It has the same covariance law as a link, but it is generally a complex
matrix rather than an element of \(SU(N)\).  A projection back to the group
is therefore part of the regulator definition.

\begin{proposition}[Equivariance of polar link projection]
\label{prop:polar-link-projection-equivariance}
Let \(X\in GL(N,\mathbb C)\), and write its polar decomposition as
\[
  X=U_XH_X,\qquad
  U_X\in U(N),\qquad H_X=(X^\dagger X)^{1/2}>0.
\]
For \(g,h\in SU(N)\),
\[
  U_{gXh^{-1}}=gU_Xh^{-1}.
  \label{eq:polar-unitary-equivariance}
\]
Consequently the locally defined special-unitary projection
\[
  \Pi_{SU(N)}(X)
  =
  (\det U_X)^{-1/N}U_X
  \label{eq:suN-polar-projection}
\]
is gauge covariant on any open set where a branch of the \(N\)-th root is
chosen continuously.
\end{proposition}

\begin{proof}
Since
\[
  (gXh^{-1})^\dagger(gXh^{-1})
  =
  h(X^\dagger X)h^{-1},
\]
functional calculus gives
\[
  H_{gXh^{-1}}=hH_Xh^{-1}.
\]
Therefore
\[
  gXh^{-1}
  =
  (gU_Xh^{-1})(hH_Xh^{-1})
\]
is the polar decomposition of \(gXh^{-1}\), because the first factor is
unitary and the second is positive.  This proves
\eqref{eq:polar-unitary-equivariance}.  The determinant of \(gU_Xh^{-1}\)
equals \(\det U_X\), since \(g,h\in SU(N)\), so the determinant-normalized
projection has the same covariance law whenever the root branch is fixed on
the chosen connected domain.
\end{proof}

The projection map is smooth on \(GL(N,\mathbb C)\) after choosing the
determinant branch, and it becomes singular when \(X\) loses rank or crosses
a branch cut.  This is why differentiable algorithms often use stout
smearing, which stays on the group without a polar projection.

Let
\[
  \Pi_{\mathfrak{su}(N)}(M)
  =
  \frac12(M-M^\dagger)
  -
  \frac{1}{2N}\operatorname{Tr}(M-M^\dagger)\mathbf 1 .
  \label{eq:suN-algebra-projection}
\]
This is the orthogonal projection, with respect to the real Hilbert--Schmidt
inner product, onto anti-Hermitian traceless matrices.  Define
\[
  \Omega_\ell(U)=C_\ell(U)U_\ell^{-1},
  \qquad
  Q_\ell(U)=\rho\,\Pi_{\mathfrak{su}(N)}(\Omega_\ell(U)),
  \qquad
  \mathcal S_\ell^{\rm stout}(U)=e^{Q_\ell(U)}U_\ell .
  \label{eq:stout-smearing-definition}
\]

\begin{proposition}[Stout smearing is a smooth gauge-covariant link map]
\label{prop:stout-smearing-gauge-covariance}
For every real \(\rho\), the map
\eqref{eq:stout-smearing-definition} is a smooth map
\(\mathcal U_\Lambda\to SU(N)^E\) and satisfies the gauge-covariance law
\eqref{eq:smearing-gauge-covariance}.
\end{proposition}

\begin{proof}
The staple covariance \eqref{eq:staple-link-covariance} gives
\[
  \Omega_\ell(g\cdot U)
  =
  g(x)\Omega_\ell(U)g(x)^{-1}.
\]
The projection \(\Pi_{\mathfrak{su}(N)}\) commutes with unitary conjugation,
so
\[
  Q_\ell(g\cdot U)=g(x)Q_\ell(U)g(x)^{-1}.
\]
Hence
\[
  e^{Q_\ell(g\cdot U)}
  =
  g(x)e^{Q_\ell(U)}g(x)^{-1}.
\]
Multiplying by the transformed link
\((g\cdot U)_\ell=g(x)U_\ell g(y)^{-1}\) gives
\[
  \mathcal S_\ell^{\rm stout}(g\cdot U)
  =
  g(x)\mathcal S_\ell^{\rm stout}(U)g(y)^{-1}.
\]
Because \(Q_\ell\) is anti-Hermitian and traceless, \(e^{Q_\ell}\in SU(N)\),
and therefore \(e^{Q_\ell}U_\ell\in SU(N)\).  Smoothness follows from the
polynomial staple map, the linear projection
\eqref{eq:suN-algebra-projection}, and the matrix exponential.
\end{proof}

Hypercubic smearing is an iterated local construction in which staples are
restricted to lower-dimensional faces before the next level is formed.  Its
definition is a finite composition of maps of the preceding type; hence its
gauge covariance follows by induction once every intermediate projection or
exponential map is specified.  Locality is regulator locality: after \(k\)
smearing steps the smeared link depends on links in a bounded neighborhood
whose radius grows at most linearly in \(k\).  A continuum theorem using
smeared operators must keep this radius fixed in lattice units, or specify
how it scales with \(a\), before the operator scheme has mathematical
meaning.

\section{Lattice Perturbative Coordinates and Tadpole Normalization}
\label{sec:lattice-perturbative-coordinates-tadpole}

Perturbation theory on a compact gauge lattice begins after choosing a chart
near the trivial connection.  For \(G=SU(N)\), use Hermitian generators
\(t^a\) normalized by
\[
  \operatorname{tr}(t^a t^b)=\delta^{ab},
  \label{eq:lattice-perturbative-trace-normalization}
\]
and write a link near the identity as
\[
  U_\mu(x)
  =
  \exp\{-\ii g_0 a A_\mu(x+\tfrac a2\widehat\mu)\},
  \qquad
  A_\mu=A_\mu^a t^a .
  \label{eq:lattice-link-perturbative-coordinate}
\]
With the Wilson action in the convention
\[
  S_W[U]
  =
  \frac{\beta}{N}
  \sum_{x,\mu<\nu}
  \operatorname{Re}\operatorname{tr}\bigl(\mathbf 1-U_{\mu\nu}(x)\bigr),
  \qquad
  \beta=\frac{N}{g_0^2},
  \label{eq:lattice-perturbative-wilson-beta}
\]
the parameter \(g_0\) is the lattice coupling coordinate matching the
continuum trace-delta convention
\(\frac{1}{4g_0^2}\int\operatorname{tr}(F_{\mu\nu}F_{\mu\nu})\) at tree
level.

Let
\[
  \nabla_\mu^+ f(x)=\frac{f(x+a\widehat\mu)-f(x)}{a}
\]
be the forward difference.  The linearized lattice curvature is
\[
  f_{\mu\nu}(x)
  =
  \nabla_\mu^+ A_\nu(x+\tfrac a2\widehat\nu)
  -
  \nabla_\nu^+ A_\mu(x+\tfrac a2\widehat\mu).
  \label{eq:lattice-linearized-curvature}
\]
Expanding the plaquette gives
\[
  U_{\mu\nu}(x)
  =
  \exp\{-\ii g_0 a^2 f_{\mu\nu}(x)+O(g_0a^3A,g_0^2a^3A^2)\}.
  \label{eq:lattice-plaquette-perturbative-expansion}
\]
Therefore the quadratic Wilson action is
\[
  S_W^{(2)}[A]
  =
  \frac12
  \sum_x a^D\sum_{\mu<\nu}
  \operatorname{tr}\bigl(f_{\mu\nu}(x)f_{\mu\nu}(x)\bigr).
  \label{eq:lattice-quadratic-wilson-action}
\]
All higher vertices come from the same compact-link exponential and from the
Baker--Campbell--Hausdorff expansion of the plaquette.  The finite-regulator
definition remains the Haar integral; the perturbative chart is a local
coordinate system for asymptotic expansion around \(U_\ell=\mathbf 1\).

On a periodic lattice, use Fourier modes with lattice momentum
\[
  \widehat p_\mu=\frac{2}{a}\sin\frac{ap_\mu}{2},
  \qquad
  \widehat p^2=\sum_\rho \widehat p_\rho^2.
  \label{eq:lattice-hat-momentum}
\]
After absorbing the harmless mid-link phases into the Fourier components,
\eqref{eq:lattice-quadratic-wilson-action} becomes
\[
  S_W^{(2)}
  =
  \frac12
  \sum_p
  A_\mu^a(-p)
  \left[
    \delta_{\mu\nu}\widehat p^2-\widehat p_\mu\widehat p_\nu
  \right]
  A_\nu^a(p).
  \label{eq:lattice-transverse-quadratic-kernel}
\]
The zero modes and gauge orbits are treated by a finite-volume gauge-fixing
or by restricting to gauge-invariant observables.  In a covariant perturbative
calculation one adds the lattice gauge-fixing quadratic form
\[
  S_{\rm gf}^{(2)}
  =
  \frac{1}{2\xi}
  \sum_p
  A_\mu^a(-p)\,\widehat p_\mu\widehat p_\nu\,A_\nu^a(p),
  \qquad \xi>0 .
  \label{eq:lattice-covariant-gauge-fixing-quadratic}
\]
The resulting kernel is
\[
  K_{\mu\nu}^{(\xi)}(p)
  =
  \delta_{\mu\nu}\widehat p^2
  -
  \left(1-\frac1\xi\right)\widehat p_\mu\widehat p_\nu,
  \label{eq:lattice-gauge-fixed-kernel}
\]
with inverse on nonzero momenta
\[
  D_{\mu\nu}^{ab}(p)
  =
  \delta^{ab}
  \frac{1}{\widehat p^2}
  \left[
    \delta_{\mu\nu}
    -
    (1-\xi)\frac{\widehat p_\mu\widehat p_\nu}{\widehat p^2}
  \right].
  \label{eq:lattice-tree-propagator}
\]
As \(a\to0\) at fixed physical \(p\),
\[
  \widehat p^2
  =
  p^2-\frac{a^2}{12}\sum_\mu p_\mu^4+O(a^4p^6).
  \label{eq:lattice-hat-momentum-expansion}
\]
This displayed artifact is one elementary source of the tree-level Symanzik
corrections discussed above.

\begin{proposition}[Tree-level lattice propagator]
\label{prop:lattice-tree-level-propagator}
For every nonzero lattice momentum, the tensor
\eqref{eq:lattice-tree-propagator} is the inverse of
\eqref{eq:lattice-gauge-fixed-kernel}:
\[
  K_{\mu\rho}^{(\xi)}(p)D_{\rho\nu}^{ab}(p)
  =
  \delta^{ab}\delta_{\mu\nu}.
  \label{eq:lattice-propagator-inverse-identity}
\]
\end{proposition}

\begin{proof}
Let \(P_L\) be the rank-one projector
\[
  (P_L)_{\mu\nu}=\frac{\widehat p_\mu\widehat p_\nu}{\widehat p^2},
  \qquad
  P_T=\mathbf 1-P_L .
\]
Then
\[
  K^{(\xi)}=\widehat p^2(P_T+\xi^{-1}P_L),
  \qquad
  D=\widehat p^{-2}(P_T+\xi P_L).
  \]
Since \(P_TP_L=0\) and \(P_T^2=P_T\), \(P_L^2=P_L\), their product is the
identity tensor.  The color factor is diagonal by
\eqref{eq:lattice-perturbative-trace-normalization}.
\end{proof}

Mean-link or tadpole normalization is best understood as a finite coordinate
change for perturbative expansions of link observables and action
coefficients.  Define the plaquette mean link, when the expectation is
positive, by
\[
  u_0(\beta)
  =
  \left\langle
    \frac1N\operatorname{Re}\operatorname{tr}U_p
  \right\rangle_\beta^{1/4}.
  \label{eq:plaquette-mean-link-definition}
\]
If a Wilson-line observable contains \(L\) elementary links, its
plaquette-tadpole-normalized coordinate is
\[
  O_{\rm TI}=u_0^{-L}O .
  \label{eq:tadpole-normalized-line-operator}
\]
For the Wilson plaquette action, the same bookkeeping rewrites the plaquette
coefficient as
\[
  \frac{\beta}{N}\operatorname{Re}\operatorname{tr}U_p
  =
  \frac{\beta u_0^4}{N}
  \operatorname{Re}\operatorname{tr}\bigl(u_0^{-4}U_p\bigr),
  \]
so the boosted coupling coordinate is
\[
  g_{\rm TI}^2=\frac{g_0^2}{u_0^4}.
  \label{eq:tadpole-boosted-coupling}
\]
Equations \eqref{eq:tadpole-normalized-line-operator} and
\eqref{eq:tadpole-boosted-coupling} define a perturbative coordinate
convention.  They acquire predictive content only after specifying which
renormalized observable is being matched and which finite-order truncation is
being compared.

\begin{example}[First terms of the boosted coupling]
\label{ex:tadpole-boosted-coupling-series}
Suppose the mean plaquette has a weak-coupling expansion
\[
  \left\langle
    \frac1N\operatorname{Re}\operatorname{tr}U_p
  \right\rangle_\beta
  =
  1-c_P g_0^2+O(g_0^4).
  \label{eq:mean-plaquette-weak-coupling}
\]
Then \(u_0^4=1-c_Pg_0^2+O(g_0^4)\) and
\[
  g_{\rm TI}^2
  =
  g_0^2\left(1+c_Pg_0^2+O(g_0^4)\right).
  \label{eq:tadpole-boosted-coupling-series}
\]
The coefficient \(c_P\) depends on the action, gauge group, and regulator
volume used to define the expectation value.  The formula itself is the
finite algebra of the coordinate change.
\end{example}

\section{Wilson Loops}

For a closed oriented lattice contour \(C\),
\[
  W_R(C)=\operatorname{Tr}_R\prod_{\ell\in C}U_\ell
\]
is gauge invariant.  Wilson loops are the basic gauge-invariant probes of
line charges.  Their strong-coupling character expansion gives the area-law
mechanism described in Chapter~\ref{ch:global-confinement-screening-oblique}.
Their continuum limit requires line renormalization and an operator scheme.

\section{Static Potential and Creutz Ratios}
\label{sec:lattice-static-potential-creutz-ratios}

A rectangular Wilson loop has a Hamiltonian meaning when the transfer matrix
exists.  Let \(C_{n,m}\) be a rectangle with spatial side \(R=na\) and
Euclidean time side \(T=ma\).  Use the normalized loop
\[
  W_S(n,m)
  =
  \frac{1}{\dim S}
  \left\langle
    \operatorname{Tr}_S\prod_{\ell\in C_{n,m}}U_\ell
  \right\rangle .
  \label{eq:normalized-rectangular-wilson-loop}
\]
In the transfer-matrix Hilbert space, the spatial transporter between two
static sources defines a vector \(\Psi_{S,n}\) in the sector containing a
source and an antisource separated by \(n\) lattice spacings.  Reflection
positivity gives
\[
  W_S(n,m)=
  \langle \Psi_{S,n},T^m\Psi_{S,n}\rangle .
  \label{eq:rectangular-wilson-loop-transfer-matrix}
\]
If the spectral measure of \(\Psi_{S,n}\) for \(T=e^{-aH}\) has an isolated
largest transfer-matrix eigenvalue \(e^{-aE_0(n,S)}\) with nonzero weight,
then
\[
  aV_{a,S}(n)
  =
  -\lim_{m\to\infty}
  \log\frac{W_S(n,m+1)}{W_S(n,m)}
  =
  aE_0(n,S).
  \label{eq:lattice-static-potential-effective-mass}
\]
This is the finite-cutoff static potential in the chosen static-source
sector.  In a continuum limit it contains an additive line self-energy:
the force, potential differences, or a specified subtraction scheme must be
used before comparing continuum quantities.

\begin{proposition}[Spectral extraction of the finite-lattice static energy]
\label{prop:spectral-static-potential-extraction}
Assume
\[
  W_m=\sum_{\alpha=0}^{r}A_\alpha e^{-amE_\alpha},
  \qquad
  A_0>0,\quad A_\alpha\ge0,\quad
  E_0<E_1\le\cdots\le E_r .
  \label{eq:finite-static-spectral-sum}
\]
Then
\[
  -\log\frac{W_{m+1}}{W_m}
  =
  aE_0+O(e^{-am(E_1-E_0)}).
  \label{eq:finite-static-effective-mass-error}
\]
\end{proposition}

\begin{proof}
Write
\[
  W_m=A_0e^{-amE_0}
  \left(
    1+\sum_{\alpha>0}\frac{A_\alpha}{A_0}
    e^{-am(E_\alpha-E_0)}
  \right).
\]
The ratio \(W_{m+1}/W_m\) is \(e^{-aE_0}\) times a quotient of two factors
which both tend to \(1\), and the difference from \(1\) is bounded by a
constant times \(e^{-am(E_1-E_0)}\).  Taking the logarithm gives
\eqref{eq:finite-static-effective-mass-error}.
\end{proof}

The Creutz ratio is a finite-difference observable designed to cancel the
leading perimeter and constant terms in a rectangular Wilson-loop ansatz.
For positive Wilson loops define
\[
  \chi(n,m)
  =
  -\log
  \frac{W(n,m)W(n-1,m-1)}
       {W(n,m-1)W(n-1,m)} .
  \label{eq:lattice-creutz-ratio}
\]
If, over a range of lattice sizes,
\[
  W(n,m)=
  \exp\{-\sigma_{\rm lat}nm-\mu_{\rm lat}(n+m)-c_{\rm lat}
       +\varepsilon(n,m)\},
  \label{eq:creutz-area-perimeter-ansatz}
\]
then
\[
  \chi(n,m)=
  \sigma_{\rm lat}
  -
  \varepsilon(n,m)-\varepsilon(n-1,m-1)
  +\varepsilon(n,m-1)+\varepsilon(n-1,m).
  \label{eq:creutz-ratio-error-combination}
\]
Thus an observed Creutz-ratio plateau is a statement about a controlled
two-dimensional finite difference of the correction term \(\varepsilon\).
It is not, by itself, a continuum string-tension theorem; that requires the
scaling trajectory, infinite-volume limit, and line-operator
renormalization to be controlled.

The companion analysis script
\[
  \texttt{qft\_scripts/static\_potential\_from\_wilson\_loops.py}
\]
implements exactly the finite formulas in
\eqref{eq:lattice-static-potential-effective-mass} and
\eqref{eq:lattice-creutz-ratio} for positive rectangular Wilson-loop data.
It reads a table of \(W(R,T)\) values, returns effective-mass ratios and
Creutz ratios, and includes a synthetic area-plus-perimeter smoke test for
\eqref{eq:creutz-area-perimeter-ansatz}.  Statistical correlations among
Monte Carlo Wilson loops belong to the upstream jackknife or bootstrap
analysis; the script's optional elementary error propagation is only the
independent-error algebra of the displayed logarithms.

\section{Wilson Flow, Smoothing, and Topological Charge}
\label{sec:lattice-wilson-flow-smoothing-topology}

Gradient flow is a finite-regulator operation before it is a continuum
renormalization tool.  Let
\[
  \mathcal U_\Lambda=G^E
\]
be the compact manifold of link variables on the finite lattice, and equip
each copy of \(G\) with the bi-invariant Riemannian metric obtained from an
invariant positive inner product \((X,Y)_{\mathfrak g}\) on
\(\mathfrak g=\operatorname{Lie}(G)\).  For a smooth gauge-invariant lattice
action \(S:\mathcal U_\Lambda\to\mathbb R\), define the left-link gradient
\(\nabla_\ell S(U)\in\mathfrak g\) by
\begin{equation}
  (X,\nabla_\ell S(U))_{\mathfrak g}
  =
  \left.\frac{\dd}{\dd s}
  S(U_{\ell'}^{(s)})\right|_{s=0},
  \qquad
  U_\ell^{(s)}=e^{sX}U_\ell,\quad U_{\ell'}^{(s)}=U_{\ell'}\ {\rm for}\ \ell'\ne\ell .
  \label{eq:lattice-link-gradient-definition}
\end{equation}
The Wilson-flow trajectory \(V_t\in\mathcal U_\Lambda\) starting at
\(U\in\mathcal U_\Lambda\) is the solution of
\begin{equation}
  \frac{\dd V_t(\ell)}{\dd t}\,V_t(\ell)^{-1}
  =
  -\nabla_\ell S(V_t),
  \qquad
  V_0(\ell)=U_\ell .
  \label{eq:lattice-wilson-flow-ode}
\end{equation}
The parameter \(t\) has engineering dimension \(a^2\) if one keeps physical
units.  The normalized flow time \(\tau=t/a^2\) is often used in finite-lattice
computations, but the continuum smoothing interpretation concerns fixed
physical \(t>0\) as \(a\to0\).

\begin{proposition}[Finite-lattice Wilson flow]
\label{prop:finite-lattice-wilson-flow}
For every finite lattice and every smooth action \(S\), the flow
\eqref{eq:lattice-wilson-flow-ode} exists for all \(t\ge0\).  It is gauge
covariant, and the action decreases according to
\begin{equation}
  \frac{\dd}{\dd t}S(V_t)
  =
  -\sum_{\ell\in E}\|\nabla_\ell S(V_t)\|_{\mathfrak g}^2
  \le0 .
  \label{eq:lattice-wilson-flow-monotonicity}
\end{equation}
\end{proposition}

\begin{proof}
The right-hand side of \eqref{eq:lattice-wilson-flow-ode} is a smooth vector
field on the compact manifold \(G^E\).  Ordinary differential-equation
existence gives a local solution, and compactness prevents finite-time escape,
so the solution is global.  Gauge covariance follows because \(S\) and the
metric are invariant under the lattice gauge group: differentiating
\(S(g\cdot U)=S(U)\) along a left variation of one link shows that the gradient
is transported by the same endpoint conjugations as the link tangent vector.
Finally, by the chain rule and
\eqref{eq:lattice-link-gradient-definition},
\[
  \frac{\dd}{\dd t}S(V_t)
  =
  \sum_{\ell\in E}
  \left(
  \frac{\dd V_t(\ell)}{\dd t}V_t(\ell)^{-1},
  \nabla_\ell S(V_t)
  \right)_{\mathfrak g}
  =
  -\sum_{\ell\in E}\|\nabla_\ell S(V_t)\|_{\mathfrak g}^2 .
\]
\end{proof}

The continuum reason that positive flow time is useful is already visible in
the linearized equation.  Let \(B_\mu(t,x)\) be a smooth continuum connection
and write its curvature as \(G_{\mu\nu}(t)\).  The Yang--Mills gradient flow is
\begin{equation}
  \partial_t B_\mu=D_\nu G_{\nu\mu},
  \qquad
  B_\mu(0,x)=A_\mu(x),
  \label{eq:continuum-gradient-flow}
\end{equation}
in the anti-Hermitian connection convention used in the lattice expansion.
Linearizing about \(A_\mu=0\) gives
\[
  \partial_t B_\mu
  =
  \partial^2B_\mu-\partial_\mu\partial_\nu B_\nu .
\]
In the transverse gauge \(\partial_\mu B_\mu=0\), a Fourier mode of momentum
\(k\) is multiplied by \(e^{-tk^2}\).  Thus positive flow time damps momenta
\(k^2\gg t^{-1}\), and the heat-kernel radius is of order \(\sqrt{8t}\) in
four Euclidean dimensions.

\begin{definition}[Flowed energy density and scale coordinates]
\label{def:lattice-flowed-energy-scale}
Let \(G_{\mu\nu}(t,x)\) be a chosen gauge-covariant lattice curvature
constructed from the flowed links \(V_t\), for example the clover curvature.
The flowed energy density is represented by
\[
  E(t,x)=-\frac12\operatorname{tr}\bigl(G_{\mu\nu}(t,x)G_{\mu\nu}(t,x)\bigr),
\]
for anti-Hermitian curvature.  Its volume average is
\[
  \overline E(t)=\frac{a^4}{V_\Lambda}\sum_x E(t,x),
  \qquad
  V_\Lambda=a^4\#\Lambda^0 .
\]
Given a positive reference number \(c\), the finite-cutoff scale coordinates
\(t_0\) and \(w_0\), if they exist uniquely, are defined by
\[
  t_0^2\,\langle\overline E(t_0)\rangle=c,
  \qquad
  \left.
  t\frac{\dd}{\dd t}\bigl(t^2\langle\overline E(t)\rangle\bigr)
  \right|_{t=w_0^2}=c .
\]
Existence, uniqueness, and a continuum limit of these coordinates are
properties of a scaling trajectory; they are not consequences of the
finite-lattice ODE alone.
\end{definition}

The same distinction is essential for topological charge.  A generic finite
lattice gauge field is a collection of group elements, not a smooth bundle
connection with a canonical integer \(Q\).  A topological sector on the lattice
requires an additional construction, such as an admissibility condition plus a
geometric interpolation, or a chiral lattice Dirac operator whose index is
well-defined.

\begin{definition}[Admissible geometric topological charge]
\label{def:admissible-geometric-topological-charge}
Let \(G=SU(N)\).  On a periodic lattice in four dimensions, an admissibility
condition is the bound
\[
  \|1-U_p\|<\epsilon
  \qquad{\rm for\ every\ plaquette}\ p
\]
with \(\epsilon\) small enough that plaquette holonomies stay in a coordinate
neighborhood of the identity.  Together with a specified interpolation of the
link data to transition functions on the continuum torus, this determines a
principal \(SU(N)\)-bundle \(P_U\to T^4\).  The geometric topological charge is
\[
  Q_{\rm geom}(U)
  =
  \frac{1}{8\pi^2}\int_{T^4}\operatorname{tr}(F_{\widetilde A}\wedge
  F_{\widetilde A})
  \in\mathbb Z,
\]
where \(\widetilde A\) is any smooth connection on \(P_U\).  Independence of
\(\widetilde A\) is the Chern--Weil theorem; independence of the interpolation
requires the stated admissibility construction and is part of the regulator
datum.
\end{definition}

\begin{proposition}[Continuum flow preserves Chern--Weil charge]
\label{prop:continuum-flow-preserves-cw-charge}
Let \(B(t)\) be a smooth one-parameter family of connections on a fixed
principal \(SU(N)\)-bundle over a closed oriented four-manifold \(M\), with
curvature \(G(t)\).  Then
\[
  Q(t)=\frac{1}{8\pi^2}\int_M\operatorname{tr}(G(t)\wedge G(t))
\]
is independent of \(t\).  In particular, smooth Yang--Mills gradient flow
cannot change the bundle topological charge unless the flow leaves the smooth
fixed-bundle setting.
\end{proposition}

\begin{proof}
Write \(\dot B=\partial_tB\).  The curvature variation is
\(\dot G=D_B\dot B\).  Therefore
\[
  \frac{\dd}{\dd t}\operatorname{tr}(G\wedge G)
  =
  2\operatorname{tr}(D_B\dot B\wedge G).
\]
Using invariance of the trace and the Bianchi identity \(D_BG=0\),
\[
  \operatorname{tr}(D_B\dot B\wedge G)
  =
  \dd\,\operatorname{tr}(\dot B\wedge G).
\]
Thus
\[
  \frac{\dd Q}{\dd t}
  =
  \frac{1}{4\pi^2}\int_M
  \dd\,\operatorname{tr}(\dot B\wedge G)
  =
  0
\]
because \(M\) is closed.
\end{proof}

\begin{controlledapproximation}[Using flow in continuum extrapolations]
\label{ca:lattice-gradient-flow-continuum-use}
A flowed observable at positive physical flow time is a composite observable
with its own regulator definition.  Statements such as ``the flow removes
ultraviolet noise'' mean the following precise finite-cutoff procedure:
first solve \eqref{eq:lattice-wilson-flow-ode}, then construct a
gauge-invariant observable from \(V_t\), and only then study its infinite-volume
and continuum limits.  Integer topological charge plateaux observed after
flow are numerical evidence for stability within an admissible sector; without
an admissibility or index-theoretic definition they are not, by themselves, a
definition of a continuum theta sector.
\end{controlledapproximation}

\section{Reflection Positivity and Transfer Matrix}

For the Wilson plaquette action, reflection positivity follows by assigning
plaquettes to the two halves of the lattice and expanding cross-plane
plaquette Boltzmann weights in positive characters.  The reflection-positive
inner product on positive-time functions gives a transfer matrix
\[
  T=e^{-aH}
\]
with \(H\geq0\) after quotienting null vectors.  Gauge constraints are
implemented by the gauge-invariant subspace or by a gauge-covariant transfer
matrix followed by projection.

\section{Fermions and Chiral Difficulties}

Adding lattice fermions requires a fermion action whose reflection
positivity, locality, spectrum, and continuum chiral properties are checked
separately.  The mid-link fermion reflection structure of
Chapter~\ref{ch:lattice-reflection-positivity} is one such finite-regulator
positivity mechanism.  Chiral gauge theories require more: gauge invariance,
locality, anomaly cancellation, and a definition of the fermion determinant
phase compatible with the target continuum theory.

\begin{openproblem}[Lattice continuum QCD construction]
\label{op:lattice-continuum-qcd-construction}
Construct four-dimensional continuum Yang--Mills or QCD from Wilson-type
lattice gauge theory with a mass gap, OS reconstruction, local gauge-invariant
operator limits, and controlled line-operator and theta-angle sectors.
\end{openproblem}
