\chapter{Hamiltonian Truncation, DLCQ, And Benchmark Protocols}
\label{ch:constructive-hamiltonian-truncation-dlcq-benchmarks}

\ConventionsLorentzian

Hamiltonian numerical methods are regulator frameworks for QFT.  This
chapter formulates Hamiltonian truncation, light-front discretization, and
benchmark scripts as controlled approximations with explicitly declared
Hilbert spaces, cutoffs, counterterms, and convergence tests.

\begin{definition}[Hamiltonian truncation regulator specification]
\label{def:hamiltonian-truncation-regulator-datum}
A Hamiltonian truncation regulator specification is a tuple
\[
  \mathfrak H_{\rm trunc}
  =
  \bigl(
    \Hilb,\mathcal Q_H,
    \{\Hilb_\Lambda,P_\Lambda\}_{\Lambda\in I},
    \mathcal B_\Lambda,
    \{\mathcal O_a,c_a(\Lambda)\}_{a\in\mathcal A_{\rm ct}},
    \mathcal O_{\rm obs},
    \mathcal E_{\rm sys}
  \bigr)
\]
with the following entries.
\begin{enumerate}[label=\textnormal{(\roman*)}]
\item \(\Hilb\) is the target Hilbert space and \(\mathcal Q_H\) is a
  lower-bounded closed quadratic form on a dense form domain
  \(\operatorname{Dom}\mathcal Q_H\).  When the limiting Hamiltonian is known
  as a self-adjoint operator \(H\), \(\mathcal Q_H\) is its associated form.
\item \(I\) is a directed cutoff set.  For each \(\Lambda\in I\),
  \(\Hilb_\Lambda\subset\operatorname{Dom}\mathcal Q_H\) is a
  finite-dimensional trial space and \(P_\Lambda\) is the Hilbert-space
  projection onto \(\Hilb_\Lambda\).  The specification states whether the intended
  density statement is strong, form-norm, graph-norm, or another declared
  topology.
\item \(\mathcal B_\Lambda\) is the ordered computational basis used to build
  finite matrices.  The basis choice includes inner-product conventions,
  quantum-number sectors, symmetry projections, and normalization of finite
  volume states.
\item \(\mathcal O_a\) are local counterterm or improvement operators admitted
  by the regulator, and \(c_a(\Lambda)\) are their cutoff-dependent
  coordinates in a declared scheme.  The finite Hamiltonian matrix is
  \[
    H_\Lambda^{\rm ren}
    =
    P_\Lambda H P_\Lambda
    +
    \sum_{a\in\mathcal A_{\rm ct}}
    c_a(\Lambda)P_\Lambda\mathcal O_aP_\Lambda
  \]
  when the formal operator expression is meaningful; otherwise the same
  formula denotes the finite quadratic form specified in the basis
  \(\mathcal B_\Lambda\).
\item \(\mathcal O_{\rm obs}\) is the list of finite-regulator observables to
  be extracted from \(H_\Lambda^{\rm ren}\), together with their normalization
  and symmetry-sector labels.
\item \(\mathcal E_{\rm sys}\) is the systematic-error model: cutoff
  extrapolation ansatz, finite-volume correction model, counterterm-fitting
  prescription, residual-bound tolerance, and benchmark scripts used to
  check finite-matrix algebra.
\end{enumerate}
This records a regulator family.  A continuum or infinite-volume claim
requires an additional convergence theorem or controlled approximation that
states the topology, the limiting quadratic form or observable, and an error
bound.
\end{definition}

\section{Projected Hamiltonian Regulator}

The elementary version of Definition~\ref{def:hamiltonian-truncation-regulator-datum}
starts with a known self-adjoint Hamiltonian \(H\) on \(\Hilb\), bounded
below.  Choose finite-dimensional subspaces
\[
  \Hilb_\Lambda\subset\Hilb,\qquad P_\Lambda:\Hilb\to\Hilb_\Lambda,
\]
with \(P_\Lambda\to I\) strongly.  The projected Hamiltonian is
\[
  H_\Lambda=P_\Lambda H P_\Lambda
  \]
on \(\Hilb_\Lambda\).  If \(H\) is defined by a QFT regulator, \(H_\Lambda\)
must be supplemented by local counterterms:
\[
  H_\Lambda^{\mathrm{ren}}
  =
  P_\Lambda H P_\Lambda
  +
  \sum_a c_a(\Lambda)P_\Lambda\mathcal O_aP_\Lambda.
\]
The pair \((\Hilb_\Lambda,H_\Lambda^{\mathrm{ren}})\) is the numerical
regulator.

\section{Rayleigh--Ritz Bound}

\begin{lemma}[Rayleigh--Ritz bound at fixed Hamiltonian]
\label{lem:rayleigh-ritz-fixed-hamiltonian}
Let \(H\) be a self-adjoint operator on \(\Hilb\), bounded below, and suppose
that
\[
  E_0\le E_1\le\cdots\le E_n
\]
are isolated eigenvalues below the bottom of the essential spectrum, counted
with multiplicity.  Let \(\mathcal Q_H\) be the closed quadratic form of
\(H\).  For a finite-dimensional subspace
\(\Hilb_\Lambda\subset\operatorname{Dom}\mathcal Q_H\), define
\[
  E_n(\Lambda)
  =
  \min_{\substack{W\subset\Hilb_\Lambda\\ \dim W=n+1}}
  \max_{\substack{\psi\in W\\ \|\psi\|=1}}\mathcal Q_H(\psi,\psi).
\]
Then \(E_n(\Lambda)\ge E_n\).  If \(\{\Hilb_\Lambda\}\) is an increasing net
and, for some orthonormal eigenbasis
\(\varphi_0,\ldots,\varphi_n\) of the first \(n+1\) eigenspaces, there are
\(\eta_j^\Lambda\in\Hilb_\Lambda\) with
\[
  \eta_j^\Lambda\to\varphi_j
  \quad\text{in the form norm}\quad
  \|\psi\|_{\mathcal Q}^2
  =
  \|\psi\|^2+\mathcal Q_H(\psi,\psi)-c\|\psi\|^2
\]
for a constant \(c<\inf\operatorname{spec}H\), then
\(E_n(\Lambda)\downarrow E_n\).
\end{lemma}

\begin{proof}
The min--max characterization of the \(n\)-th discrete eigenvalue gives
\[
  E_n
  =
  \inf_{\substack{W\subset\operatorname{Dom}\mathcal Q_H\\ \dim W=n+1}}
  \sup_{\substack{\psi\in W\\ \|\psi\|=1}}\mathcal Q_H(\psi,\psi).
\]
The variational problem defining \(E_n(\Lambda)\) restricts the same infimum
to the smaller family of subspaces contained in \(\Hilb_\Lambda\), hence
\(E_n(\Lambda)\ge E_n\).  By the approximation hypothesis, Gram--Schmidt in
the ordinary Hilbert norm produces an \((n+1)\)-dimensional subspace
\(W_\Lambda\subset\Hilb_\Lambda\) whose unit sphere converges in form norm to
the unit sphere of
\(\operatorname{span}\{\varphi_0,\ldots,\varphi_n\}\).  The Gram--Schmidt
coefficients are rational functions of inner products whose limiting
denominators are nonzero, so the form-norm convergence is preserved.  The
supremum of \(\mathcal Q_H\) on that unit sphere therefore converges to
\(E_n\).
The upper bound \(E_n(\Lambda)\le \sup_{\psi\in W_\Lambda,\|\psi\|=1}
\mathcal Q_H(\psi,\psi)\) gives \(\limsup_\Lambda E_n(\Lambda)\le E_n\).
Together with the lower bound this proves convergence.  Monotonic decrease
follows because increasing \(\Hilb_\Lambda\) enlarges the family of trial
subspaces.
\end{proof}

This proposition concerns a fixed self-adjoint operator \(H\).  It does not
assert monotonicity for a family \(H_\Lambda^{\mathrm{ren}}\) whose
counterterms change with \(\Lambda\), nor for a sequence of finite matrices
obtained from different regulators.  Those comparisons require their own
cutoff-error estimates.

\begin{proposition}[Gap-stable Ritz counting for a fixed Hamiltonian]
\label{prop:gap-stable-ritz-counting-fixed-hamiltonian}
Let \(H\) and \(\mathcal Q_H\) be as in
Lemma~\ref{lem:rayleigh-ritz-fixed-hamiltonian}.  Let
\(\{\Hilb_\Lambda\}\) be an increasing net of finite-dimensional subspaces of
\(\operatorname{Dom}\mathcal Q_H\) which is dense in the form norm.  Let
\[
  E_0\leq E_1\leq\cdots
\]
be the discrete eigenvalues of \(H\) below the bottom of the essential
spectrum, counted with multiplicity, and let
\[
  E_0(\Lambda)\leq E_1(\Lambda)\leq\cdots
       \leq E_{\dim\Hilb_\Lambda-1}(\Lambda)
\]
be the Ritz eigenvalues obtained by restricting \(\mathcal Q_H\) to
\(\Hilb_\Lambda\).  Suppose \(a<b<\inf\operatorname{spec}_{\rm ess}H\) and
that \(a,b\notin\operatorname{spec}H\).  Define
\[
  N_H(c)=\#\{j:E_j<c\},
  \qquad c\in\{a,b\},
\]
and \(m=N_H(b)-N_H(a)\).  Then, for all sufficiently large \(\Lambda\),
exactly \(m\) Ritz eigenvalues lie in the interval \((a,b)\).
\end{proposition}

\begin{proof}
Since \(b\) lies below the essential spectrum and is not an eigenvalue, there
are only finitely many eigenvalues below \(b\).  Put
\[
  M=N_H(b),\qquad r=N_H(a).
\]
For each \(j<M\), the form-norm density of \(\bigcup_\Lambda\Hilb_\Lambda\)
implies the approximation hypothesis in
Lemma~\ref{lem:rayleigh-ritz-fixed-hamiltonian} for the first
\(j+1\) eigenspaces; hence
\begin{equation}
  E_j(\Lambda)\downarrow E_j .
  \label{eq:fixed-H-ritz-eigenvalue-convergence}
\end{equation}
If \(j<r\), then \(E_j<a\), and
\eqref{eq:fixed-H-ritz-eigenvalue-convergence} gives
\[
  E_j(\Lambda)<a
\]
for all sufficiently large \(\Lambda\).  If \(r\leq j<M\), then
\[
  a<E_j<b .
\]
The lower inequality \(E_j(\Lambda)\geq E_j>a\) follows from
Lemma~\ref{lem:rayleigh-ritz-fixed-hamiltonian}; convergence gives
\[
  E_j(\Lambda)<b
\]
for all sufficiently large \(\Lambda\).  Thus, eventually, at least
\(M-r=m\) Ritz values lie in \((a,b)\), namely those with labels
\[
  r,\ldots,M-1 .
\]
It remains to exclude extra Ritz values below \(b\).  If
\(E_M(\Lambda)<b\) for some \(\Lambda\), then the min--max definition of
\(E_M(\Lambda)\) supplies an \((M+1)\)-dimensional subspace of
\(\operatorname{Dom}\mathcal Q_H\) on whose unit sphere the quadratic form is
strictly less than \(b\).  Applying the min--max characterization in the full
form domain would give \(E_M<b\), contradicting \(M=N_H(b)\).  Therefore no
Ritz value with label \(j\geq M\) can lie below \(b\).  Combining the two
statements gives exactly \(m\) Ritz values in \((a,b)\) for all sufficiently
large \(\Lambda\).
\end{proof}

The proposition is a theorem about Galerkin approximation to one fixed
self-adjoint Hamiltonian and one fixed quadratic form.  Its hypotheses are
visible: a lower-bounded self-adjoint operator, an isolated finite spectral
window, and form-norm density of the trial spaces.  It does not justify a
counterterm-dependent truncation unless one has first proved that the
renormalized finite matrices converge to a fixed limiting quadratic form in a
sense strong enough to replace the form-norm density assumption.

\section{Truncated Conformal Space}

For a two-dimensional QFT built as a relevant deformation of a CFT, choose the
CFT Hilbert space on a circle and keep states satisfying
\[
  L_0+\bar L_0-\frac{c}{12}\le E_{\mathrm{cut}}\,\frac{L}{2\pi}.
\]
The Hamiltonian matrix elements are
\[
  \langle i|H|j\rangle
  =
  \frac{2\pi}{L}
  \left(\Delta_i-\frac{c}{12}\right)\delta_{ij}
  +
  \sum_a\lambda_a
  \left(\frac{L}{2\pi}\right)^{1-\Delta_a}
  \langle i|\mathcal O_a(1)|j\rangle.
\]
High-energy states generate cutoff dependence.  The local Hamiltonian
counterterms are determined by matching low-energy observables as
\(E_{\mathrm{cut}}\) is varied.

\section{Exactly Solvable TCSA Benchmark: Ising Energy Deformation}

A useful benchmark must distinguish the finite Hamiltonian matrix from the
continuum theory it approximates.  The thermal deformation of the two
dimensional Ising CFT is special because the deformed theory is a free massive
Majorana fermion.  On a circle of circumference \(L\), in the
Neveu--Schwarz sector, positive modes are labelled by
\[
  r\in \mathbb Z_{\geq0}+\frac12,
  \qquad
  p_r=\frac{2\pi r}{L}.
\]
For each \(r\), the finite-mode Hamiltonian reduces to the real symmetric
Bogoliubov block
\[
  H_r=
  \begin{pmatrix}
    -p_r&m\\
    m&p_r
  \end{pmatrix},
\]
where \(m\) is the Majorana mass induced by the energy-operator coupling in a
fixed normalization.  The corresponding finite truncation with \(N\) modes is
the direct sum
\[
  H_N=\bigoplus_{r=1/2}^{N-1/2} H_r .
\]

\begin{remark}[Finite-mode Majorana benchmark]
\label{rem:ising-energy-tcsa-benchmark}
The eigenvalues of \(H_N\) are
\[
  \left\{
    \pm\sqrt{p_r^2+m^2}\ :\
    r=\frac12,\frac32,\ldots,N-\frac12
  \right\}.
\]
For a fixed mode \(r\),
\[
  \det(H_r-\lambda I)
  =
  (-p_r-\lambda)(p_r-\lambda)-m^2
  =
  \lambda^2-(p_r^2+m^2).
\]
Thus the two eigenvalues of \(H_r\) are
\(\lambda=\pm\sqrt{p_r^2+m^2}\).  The matrix \(H_N\) is a direct sum of the
blocks \(H_r\), so its characteristic polynomial is the product of the block
characteristic polynomials.  The benchmark is therefore exact at every finite
cutoff \(N\); any numerical deviation from these values measures an
implementation or diagonalization error rather than a truncation effect.
\end{remark}

The script
\[
  \texttt{qft\_scripts/tcsa\_ising\_energy\_benchmark.py}
\]
diagonalizes \(H_N\) and checks Remark
\ref{rem:ising-energy-tcsa-benchmark}.  Its role is narrow and rigorous: it
is a public finite-matrix benchmark for Hamiltonian-truncation conventions
before nonintegrable perturbations, counterterm fitting, and cutoff
extrapolation are introduced.  The magnetic-field Ising TCSA problem and the
\(E_8\) mass-spectrum computation require additional sectors and perturbing
operators.

\section{Sine-Gordon Zero-Mode Truncation As A Vertex Benchmark}
\label{sec:sine-gordon-zero-mode-truncation-benchmark}

The sine-Gordon model is a useful Hamiltonian-truncation test because the
perturbing operator is a pair of vertex operators.  A full finite-volume
sine-Gordon TCSA computation requires the compact-boson Hilbert space with
oscillator descendants, winding sectors, vertex-operator normal ordering,
topological-charge choices, cutoff counterterms, and an extrapolation
scheme.  Before those ingredients are assembled, there is a small finite
benchmark that checks the zero-mode vertex selection rule without pretending
to compute the complete QFT spectrum.

Let \(\theta\) be a \(2\pi\)-periodic coordinate and let
\[
  \Hilb_0=L^2(S^1,\dd\theta/2\pi),
  \qquad
  \ket n(\theta)=e^{\ii n\theta},
  \qquad n\in\mathbb Z .
\]
The zero-mode coordinate is related to the sine-Gordon field by
\(\theta=\beta q\), where \(q\) is the spatial zero mode of \(\phi\).  All
normalization factors depending on \(L\), the compactification radius, and
\(\beta\) are absorbed below into two positive finite-regulator coordinates:
\(\kappa>0\), the coefficient of the zero-mode kinetic energy, and
\(g\in\mathbb R\), the coefficient of the vertex perturbation.  For
\(N\in\mathbb Z_{\geq0}\), let
\[
  P_N:\Hilb_0\to
  \operatorname{span}\{\ket n:-N\leq n\leq N\}
\]
be the Fourier cutoff.  The finite matrix is
\begin{equation}
  H_N^{\rm SG,0}
  =
  P_N\left(-\kappa\frac{\dd^2}{\dd\theta^2}+g\cos\theta\right)P_N .
  \label{eq:sine-gordon-zero-mode-finite-matrix}
\end{equation}
In the ordered basis \(\ket{-N},\ldots,\ket N\),
\begin{equation}
  \bigl(H_N^{\rm SG,0}\bigr)_{nm}
  =
  \kappa n^2\delta_{nm}
  +
  \frac g2\left(\delta_{n,m+1}+\delta_{n,m-1}\right),
  \qquad -N\leq n,m\leq N .
  \label{eq:sine-gordon-zero-mode-tridiagonal}
\end{equation}

\begin{remark}[Benchmark scope]
\label{rem:sine-gordon-zero-mode-benchmark-scope}
The matrix \eqref{eq:sine-gordon-zero-mode-tridiagonal} checks three finite
facts that are sensitive to conventions.  First,
\(\cos\theta=(e^{\ii\theta}+e^{-\ii\theta})/2\) shifts the compact momentum
by \(\pm1\) and has no diagonal zero-mode matrix element.  Second, the
projected finite matrix is Hermitian.  Third, if \(N\geq1\), ordinary
finite-dimensional perturbation theory around the nondegenerate state
\(\ket0\) gives the second-order ground-state shift
\begin{equation}
  \Delta E_0^{(2)}
  =
  \sum_{n\neq0}
  \frac{|\bra n\,g\cos\theta\,\ket0|^2}{0-\kappa n^2}
  =
  -\frac{g^2}{2\kappa}.
  \label{eq:sine-gordon-zero-mode-second-order-shift}
\end{equation}
Only \(n=\pm1\) contributes.  None of these finite identities asserts the
existence or accuracy of a full sine-Gordon TCSA spectrum.  The full
two-dimensional calculation must add oscillator states and all regulator data
listed at the beginning of this section.
\end{remark}

The companion script
\[
  \texttt{qft\_scripts/sine\_gordon\_zero\_mode\_truncation.py}
\]
builds \eqref{eq:sine-gordon-zero-mode-tridiagonal}, diagonalizes it at a
declared finite \(N\), checks Hermiticity and reflection symmetry
\(n\mapsto -n\), and reports the perturbative coefficient
\eqref{eq:sine-gordon-zero-mode-second-order-shift}.  The calculation check
\[
  \texttt{calculation-checks/hamiltonian\_truncation\_dlcq\_checks.py}
\]
verifies the \(\Delta n=\pm1\) selection rule, the zero-coupling spectrum,
finite Hermiticity, and the second-order shift directly from the finite
matrix elements.

\section{Sine-Gordon TCSA Vertex Assembly}
\label{sec:sine-gordon-compact-boson-tcsa-vertex-assembly}

The zero-mode benchmark checks only the compact momentum shift.  A genuine
TCSA matrix for the sine-Gordon perturbation also needs the oscillator
descendants of the compact boson.  This section records the finite algebraic
object that must be assembled before any discussion of spectra, counterterms,
or extrapolation can begin.

Fix a compact free-boson radius coordinate \(R>0\).  The finite regulator
uses integer momentum and winding labels
\[
  n\in\{-N_{\rm mom},\ldots,N_{\rm mom}\},
  \qquad
  w\in\{-N_{\rm wind},\ldots,N_{\rm wind}\},
\]
left and right oscillator occupation numbers
\[
  {\bf m}=(m_1,\ldots,m_M),\qquad
  \bar{\bf m}=(\bar m_1,\ldots,\bar m_M),
  \qquad m_r,\bar m_r\in\mathbb Z_{\geq0},
\]
and the total oscillator-level cutoff
\[
  |{\bf m}|_{\rm osc}+|\bar{\bf m}|_{\rm osc}
  =
  \sum_{r=1}^{M}r(m_r+\bar m_r)
  \leq L_{\rm osc}.
\]
The finite basis vector is denoted
\(\ket{n,w;{\bf m},\bar{\bf m}}\).  In the convention used by the companion
script, its dimensionless free cylinder energy and spin are
\begin{align}
  E_0(n,w;{\bf m},\bar{\bf m})
  &=
  \frac12\left(\frac nR\right)^2
  +
  \frac12(wR)^2
  +
  |{\bf m}|_{\rm osc}
  +
  |\bar{\bf m}|_{\rm osc},
  \label{eq:sine-gordon-tcsa-free-energy-coordinate}
  \\
  S(n,w;{\bf m},\bar{\bf m})
  &=
  nw+|{\bf m}|_{\rm osc}-|\bar{\bf m}|_{\rm osc}.
  \label{eq:sine-gordon-tcsa-spin-coordinate}
\end{align}
The additive Casimir energy is irrelevant for the finite vertex-normalization
check and has been omitted.  If the calculation is performed in the spatial
momentum-zero sector, one keeps \(S=0\); equivalently, the spatial integral
of a spin-zero perturbing operator kills matrix elements whose bra and ket
spins differ.

Let \(b_r,\bar b_r\) be normalized oscillator annihilation operators,
\([b_r,b_s^\dagger]=[\bar b_r,\bar b_s^\dagger]=\delta_{rs}\).  For a real
finite vertex charge coordinate \(\alpha\), define the normal-ordered
oscillator factor
\begin{equation}
  \mathcal E_r(\alpha)
  =
  \exp\!\left(\frac{\alpha}{\sqrt r}b_r^\dagger\right)
  \exp\!\left(-\frac{\alpha}{\sqrt r}b_r\right),
  \qquad
  \bar{\mathcal E}_r(\alpha)
  =
  \exp\!\left(\frac{\alpha}{\sqrt r}\bar b_r^\dagger\right)
  \exp\!\left(-\frac{\alpha}{\sqrt r}\bar b_r\right).
  \label{eq:sine-gordon-tcsa-normal-ordered-factor}
\end{equation}
The finite matrix element of \(\mathcal E_r(\alpha)\) between normalized
number states is the finite sum
\begin{equation}
\begin{split}
  \bra{a}\mathcal E_r(\alpha)\ket{b}
  =
  \sum_{\ell=0}^{b}
  {\bf 1}_{a-b+\ell\geq0}\,
  \frac{(-s)^\ell}{\ell!}
  \sqrt{\frac{b!}{(b-\ell)!}}\,
  \frac{s^{a-b+\ell}}{(a-b+\ell)!}
  \sqrt{\frac{a!}{(b-\ell)!}},
  \qquad
  s=\frac{\alpha}{\sqrt r}.
\end{split}
  \label{eq:sine-gordon-tcsa-one-oscillator-matrix-element}
\end{equation}
The formula is an exact finite matrix element of the normal-ordered
exponential.  The sum is finite because the annihilation exponential can
remove at most \(b\) quanta, and the subsequent creation exponential must
create exactly \(a-b+\ell\) quanta.

The finite local vertex shifting the compact momentum by
\(\varepsilon=\pm1\) has the matrix element
\begin{equation}
\begin{split}
  &\bra{n',w';{\bf a},\bar{\bf a}}\,
  V_\varepsilon(0)\,
  \ket{n,w;{\bf b},\bar{\bf b}}
  \\
  &\quad =
  \delta_{n',n+\varepsilon}\,
  \delta_{w',w}\,
  \prod_{r=1}^{M}
  \bra{a_r}\mathcal E_r(\varepsilon\alpha)\ket{b_r}\,
  \prod_{r=1}^{M}
  \bra{\bar a_r}\bar{\mathcal E}_r(\varepsilon\alpha)\ket{\bar b_r}.
\end{split}
  \label{eq:sine-gordon-tcsa-local-vertex-matrix}
\end{equation}
The spatially integrated sine-Gordon perturbation in this finite convention
is
\begin{equation}
  V_{\rm SG}^{\rm fin}
  =
  \frac12\,
  \Pi_{\Delta S=0}
  \left(V_{+}(0)+V_{-}(0)\right),
  \label{eq:sine-gordon-tcsa-integrated-vertex}
\end{equation}
where \(\Pi_{\Delta S=0}\) means that the matrix element is set to zero
unless the spin coordinate \eqref{eq:sine-gordon-tcsa-spin-coordinate} is the
same on the bra and ket.  The finite TCSA matrix assembled by the script is
\begin{equation}
  H_{\rm SG}^{\rm fin}
  =
  E_0+\lambda V_{\rm SG}^{\rm fin}.
  \label{eq:sine-gordon-tcsa-finite-hamiltonian}
\end{equation}

\begin{remark}[What this finite construction verifies]
\label{rem:sine-gordon-tcsa-vertex-assembly-scope}
Equations
\eqref{eq:sine-gordon-tcsa-free-energy-coordinate}--\eqref{eq:sine-gordon-tcsa-finite-hamiltonian}
verify the finite matrix-assembly problem for the compact-boson vertex
operator.  For example, between oscillator vacua one recovers
\(\bra{n+1,w;0,0}V_{\rm SG}^{\rm fin}\ket{n,w;0,0}=1/2\).  Between a
vacuum ket and a bra with one left and one right oscillator in the first
mode, the integrated matrix element is \(\alpha^2/2\), while a bra with only
one left oscillator has zero integrated matrix element because the spin
changes.  The winding label is preserved.  These are finite algebraic
checks.  A continuum sine-Gordon spectrum also requires the physical
normalization connecting \(\alpha\) to the usual coupling \(\beta\), the
choice of radius and topological sectors, relevant counterterms generated by
the cutoff, and a controlled \(L_{\rm osc},N_{\rm mom},M\to\infty\)
extrapolation.
\end{remark}

The companion script
\[
  \texttt{qft\_scripts/sine\_gordon\_tcsa\_vertex.py}
\]
constructs the basis above, assembles
\eqref{eq:sine-gordon-tcsa-integrated-vertex}, checks Hermiticity, verifies
spin and winding selection, and diagonalizes the resulting finite matrix for
a declared cutoff.  The calculation-check harness verifies the displayed
oscillator-vacuum, one-oscillator, and two-oscillator matrix elements
directly from \eqref{eq:sine-gordon-tcsa-one-oscillator-matrix-element}.

\section{\texorpdfstring{Nonintegrable Scalar \(\phi^4\) Truncation}{Nonintegrable Scalar phi4 Truncation}}
\label{sec:finite-phi4-hamiltonian-truncation-benchmark}

The previous benchmark is exactly solvable.  It is useful for checking mode
normalizations, but it does not test the matrix assembly problem that appears
in a genuinely interacting truncation.  The simplest finite QFT benchmark is
the massive scalar field on a spatial circle, with the quartic interaction
normal ordered relative to the massive free basis.  This is still only a
finite regulator; the continuum two-dimensional \(\phi^4\) theory requires
the counterterm and cutoff analysis developed in the constructive chapters.

\begin{construction}[Finite normal-ordered scalar truncation]
\label{con:finite-normal-ordered-phi4-truncation}
Fix \(L>0\), \(m>0\), a Fourier cutoff \(N\in\mathbb Z_{\geq0}\), a particle
cutoff \(P_{\max}\), a free-energy cutoff \(E_{\rm cut}\), and an integer
total momentum label \(P_{\rm tot}\).  For
\[
  r\in\{-N,-N+1,\ldots,N\},
  \qquad
  k_r=\frac{2\pi r}{L},
  \qquad
  \omega_r=\sqrt{m^2+k_r^2},
\]
let \(a_r,a_r^\dagger\) be finite-mode creation and annihilation operators
with \([a_r,a_s^\dagger]=\delta_{rs}\).  The truncated basis is the finite
set of occupation vectors \({\bf n}=(n_{-N},\ldots,n_N)\) satisfying
\begin{equation}
  \sum_{|r|\leq N}n_r\leq P_{\max},
  \qquad
  \sum_{|r|\leq N}r n_r=P_{\rm tot},
  \qquad
  \sum_{|r|\leq N}\omega_r n_r\leq E_{\rm cut}.
  \label{eq:finite-phi4-truncation-basis}
\end{equation}
The corresponding normalized Fock vector is
\[
  \ket{\bf n}
  =
  \prod_{|r|\leq N}\frac{(a_r^\dagger)^{n_r}}{\sqrt{n_r!}}\ket{0}.
\]
The free Hamiltonian matrix is diagonal:
\[
  (H_0)_{{\bf n}{\bf m}}
  =
  \delta_{{\bf n}{\bf m}}
  \sum_{|r|\leq N}\omega_r n_r .
\]
Define the finite free field at time zero by
\[
  \phi_N(x)
  =
  \sum_{|r|\leq N}
  \frac{1}{\sqrt{2L\omega_r}}
  \left(a_r e^{\ii k_r x}+a_r^\dagger e^{-\ii k_r x}\right),
\]
and normal order with respect to the free vacuum.  The finite interaction
matrix is
\begin{equation}
  V_N
  =
  P_{\Lambda}
  \int_0^L:\phi_N(x)^4:\,\dd x\,
  P_{\Lambda},
  \label{eq:finite-phi4-normal-ordered-interaction}
\end{equation}
where \(P_\Lambda\) is the orthogonal projection onto the span of
\eqref{eq:finite-phi4-truncation-basis}.  Equivalently,
\begin{equation}
  \int_0^L:\phi_N(x)^4:\,\dd x
  =
  \sum_{\substack{s_i=\pm1,\ |r_i|\leq N\\
        \sum_i s_i r_i=0}}
  L\prod_{i=1}^4\frac{1}{\sqrt{2L\omega_{r_i}}}
  :B_{s_1,r_1}B_{s_2,r_2}B_{s_3,r_3}B_{s_4,r_4}: ,
  \label{eq:finite-phi4-mode-expanded-interaction}
\end{equation}
with \(B_{+1,r}=a_r\) and \(B_{-1,r}=a_r^\dagger\).  The Kronecker
condition in \eqref{eq:finite-phi4-mode-expanded-interaction} is precisely
the integral over the circle.  The finite benchmark Hamiltonian is
\begin{equation}
  H_\Lambda^{(\phi^4)}
  =
  H_0+\frac{\lambda}{4!}V_N .
  \label{eq:finite-phi4-truncation-hamiltonian}
\end{equation}
No continuum statement is attached to
\eqref{eq:finite-phi4-truncation-hamiltonian} unless a counterterm scheme and
a cutoff-removal theorem or controlled approximation have also been stated.
\end{construction}

The matrix \(H_\Lambda^{(\phi^4)}\) is Hermitian at each finite cutoff.  The
field \(\phi_N(x)\) is self-adjoint on the finite Fock space generated by the
retained modes, normal ordering preserves the adjoint of the polynomial, and
\(P_\Lambda\) is an orthogonal projection.  In the explicit sum
\eqref{eq:finite-phi4-mode-expanded-interaction}, taking the adjoint changes
each \(s_i\) to \(-s_i\) and reverses the operator order before normal
ordering; the coefficient is real and the constraint
\(\sum_i s_i r_i=0\) is invariant.  Thus the restricted matrix is equal to
its adjoint.  This finite Hermiticity is a minimal consistency check; it does
not estimate the dependence on \(N\), \(P_{\max}\), or \(E_{\rm cut}\).

A useful normalization check comes from retaining only the zero mode.  If
\(N=0\), then \(\omega_0=m\) and
\[
  \int_0^L:\phi_0(x)^4:\,\dd x
  =
  \frac{1}{4Lm^2}:(a_0+a_0^\dagger)^4: .
\]
The vacuum diagonal entry is zero.  In the two-particle zero-mode state
\(\ket{2}=(a_0^\dagger)^2\ket0/\sqrt2\), the only diagonal monomial is
\(6(a_0^\dagger)^2a_0^2\), hence
\begin{equation}
  \bra{2}\int_0^L:\phi_0(x)^4:\,\dd x\ket{2}
  =
  \frac{6}{4Lm^2}\bra2(a_0^\dagger)^2a_0^2\ket2
  =
  \frac{3}{Lm^2}.
  \label{eq:finite-phi4-zero-mode-two-particle-check}
\end{equation}
This is an elementary finite-regulator identity, but it is convention
sensitive: changing the field normalization, the \(4!\) convention, or the
normal-ordering prescription changes the displayed number.

The script
\[
  \texttt{qft\_scripts/phi4\_hamiltonian\_truncation.py}
\]
implements Construction~\ref{con:finite-normal-ordered-phi4-truncation}.
Its smoke mode builds a small total-momentum-zero matrix, checks Hermiticity,
and prints the lowest finite eigenvalues.  The calculation check
\[
  \texttt{calculation-checks/hamiltonian\_truncation\_dlcq\_checks.py}
\]
verifies the momentum-sector condition, the free-energy cutoff condition, the
zero-coupling reduction to \(H_0\), and the zero-mode identity
\eqref{eq:finite-phi4-zero-mode-two-particle-check}.  These checks are not a
substitute for the constructive \(\phi^4_2\) continuum theorem; they verify
the finite matrix that a reader can diagonalize before studying cutoff
renormalization.

\section{Ising TCSA And TFFSA With Spin Deformation}

The two-dimensional Ising field theory with both thermal and magnetic
deformations is the basic test case in which Hamiltonian truncation is useful
and conceptually delicate.  Start from the critical Ising CFT on a spatial
circle of circumference \(L\).  Let \(H_{\rm CFT}(L)\) be the self-adjoint
Hamiltonian generated by cylinder time translations,
\[
  H_{\rm CFT}(L)
  =
  \frac{2\pi}{L}
  \left(L_0+\bar L_0-\frac{c}{12}\right),
  \qquad c=\frac12 .
\]
Let \(\varepsilon\) be the energy operator of scaling dimension
\(\Delta_\varepsilon=1\), and let \(\sigma\) be the spin operator of scaling
dimension \(\Delta_\sigma=1/8\).  A TCSA regulator for the two-coupling Ising
family is the finite matrix
\begin{equation}
  H_{\rm TCSA}(L;E_{\rm cut})
  =
  P_{E_{\rm cut}}
  \left[
    H_{\rm CFT}(L)
    +
    \tau\int_0^L\varepsilon_{\rm cyl}(0,x)\,\dd x
    +
    h\int_0^L\sigma_{\rm cyl}(0,x)\,\dd x
  \right]
  P_{E_{\rm cut}}
  +
  H_{\rm ct}(E_{\rm cut}).
  \label{eq:ising-two-coupling-tcsa-hamiltonian}
\end{equation}
Here \(P_{E_{\rm cut}}\) projects onto CFT states satisfying
\((2\pi/L)(L_0+\bar L_0-c/12)\le E_{\rm cut}\).  The symbol
\(\mathcal O_{\rm cyl}\) means the cylinder field obtained from the plane
field by the conformal map.  After fixing a plane normalization, the
finite-volume matrix elements contain the factors \((2\pi/L)^{\Delta_a}\)
dictated by the state-operator correspondence.  The couplings have
engineering dimensions
\[
  [\tau]=1,\qquad [h]=\frac{15}{8}.
\]
The counterterm matrix \(H_{\rm ct}(E_{\rm cut})\) is part of the regulator.
It is determined by a declared matching prescription; omitting it is a choice
of truncation scheme, not a theorem that no high-energy effects exist.

The same physical family admits a second truncation.  Treat the thermal
deformation exactly.  For \(\tau\ne0\), the unperturbed theory is the free
massive Majorana theory with mass \(m\) proportional to \(\tau\) in the chosen
normalization.  The magnetic perturbation is then
\begin{equation}
  H_{\rm IFT}(L)
  =
  H_{\rm Maj}(m,L)
  +
  h\int_0^L\sigma(0,x)\,\dd x .
  \label{eq:ising-tffsa-hamiltonian}
\end{equation}
Truncating \eqref{eq:ising-tffsa-hamiltonian} in the finite-volume fermion
Fock space is the truncated free-fermion-space approach.  Its advantage is
not a new definition of the continuum theory; rather, it moves the relevant
thermal deformation from the perturbation into the basis.  This can reduce
cutoff sensitivity when the low-energy states are better approximated by
massive fermion states than by a small number of conformal states.

We now spell out the finite matrix.  Let \(\mathfrak s\) denote a fermion
spin-structure sector on the circle, with allowed momenta
\[
  p_n=\frac{2\pi n}{L},\qquad n\in Q_{\mathfrak s},
\]
where \(Q_{\mathfrak s}\) is either \(\mathbb Z+\frac12\) or
\(\mathbb Z\), depending on the sector.  A finite-volume rapidity
\(\theta_n\) is defined by
\[
  p_n=m\sinh\theta_n,\qquad
  \omega_n=m\cosh\theta_n .
\]
For a finite set \(I=\{n_1,\ldots,n_k\}\subset Q_{\mathfrak s}\), write
\[
  \ket{I}_{\mathfrak s}
  =
  a_{n_1}^\dagger\cdots a_{n_k}^\dagger\ket{0}_{\mathfrak s},
  \qquad
  E_I^{(0)}(L)=E_{\mathfrak s}^{\rm vac}(L)+\sum_{n\in I}\omega_n,
  \qquad
  P_I=\sum_{n\in I}p_n .
\]
The TFFSA cutoff is an energy cutoff in this massive basis:
\[
  \Hilb_{\rm FF}(L,E_{\rm cut})
  =
  \operatorname{span}\{\ket{I}_{\mathfrak s}:E_I^{(0)}(L)\le E_{\rm cut}\}.
\]
The diagonal part of \(H_{\rm Maj}(m,L)\) is exact in this basis.  The
nontrivial matrix is the spin perturbation.

\begin{definition}[TFFSA spin matrix element]
\label{def:tffsa-spin-matrix-element}
For two finite-volume fermion states \(I,J\), the connected infinite-volume
asymptotic spin matrix element is
\begin{equation}
  V_{IJ}^{\rm conn}
  =
  hL\,\delta_{P_I,P_J}\,
  \frac{
    F_{\rm conn}^{\sigma}(\{\theta_i:i\in I\}\mid \{\theta_j:j\in J\})
  }{
    \sqrt{\rho_I\rho_J}
  },
  \qquad
  \rho_I=\prod_{i\in I}mL\cosh\theta_i .
  \label{eq:tffsa-connected-spin-matrix-element}
\end{equation}
Here \(F_{\rm conn}^{\sigma}\) is the crossed spin-field form factor derived
in Chapter~\ref{ch:integrable-form-factor-bootstrap-local-operators}, and
\(\rho_I\) is the free Bethe-state density of states.  The full finite-volume
matrix element also contains disconnected contraction terms when \(I\) and
\(J\) share rapidities, and exponentially small finite-size corrections of
order \(e^{-mL}\) in a massive theory.  A numerical TFFSA scheme must specify
whether it uses exact finite-volume Ising form factors, the asymptotic formula
\eqref{eq:tffsa-connected-spin-matrix-element}, or a hybrid prescription.
\end{definition}

This definition makes the status of the computation explicit.  At fixed
\((L,E_{\rm cut})\), after all matrix elements have been declared, TFFSA is
ordinary finite-dimensional spectral theory.  The physical assertions concern
the removal or extrapolation of the cutoff and, if
\eqref{eq:tffsa-connected-spin-matrix-element} is used, the large-\(mL\)
control of the finite-volume form-factor approximation.

\begin{definition}[Zero-momentum connected TFFSA block]
\label{def:zero-momentum-connected-tffsa-block}
Fix \(m>0\), \(L>0\), an integer \(N_{\rm pair}\geq0\), a spin-field
normalization \(\bar\sigma\), and a magnetic coupling \(h\).  For
\[
  r_n=n+\frac12,\qquad
  p_n=\frac{2\pi r_n}{L},\qquad
  p_n=m\sinh\theta_n,\qquad
  \omega_n=m\cosh\theta_n ,
  \quad n=0,\ldots,N_{\rm pair}-1,
\]
let
\[
  \ket{0},\qquad
  \ket{n}:=\ket{\theta_n,-\theta_n},\quad
  n=0,\ldots,N_{\rm pair}-1,
\]
be the finite zero-total-momentum basis consisting of the vacuum and
two-particle pairs.  Set
\[
  \rho_n=(mL\cosh\theta_n)^2 .
\]
The connected-block TFFSA matrix is the Hermitian
\((N_{\rm pair}+1)\times(N_{\rm pair}+1)\) matrix
\[
  H_{ij}=E_i^{(0)}\delta_{ij}+hL\,V_{ij},
\]
where \(E_0^{(0)}=0\), \(E_n^{(0)}=2\omega_n\) for \(n\geq1\), and
\[
  V_{ii}=\bar\sigma .
\]
For \(i<j\), \(V_{ij}\) is defined by the crossed connected form factor
\begin{equation}
  V_{ij}
  =
  \frac{
    F_{\rm conn}^{\sigma}(
      \{\theta_a+\ii\pi:a\in I_i\}\mid \{\theta_b:b\in I_j\})
  }{
    \sqrt{\rho_i\rho_j}
  },
  \qquad
  V_{ji}=\overline{V_{ij}},
  \label{eq:zero-momentum-tffsa-block}
\end{equation}
with \(\rho_0=1\) and \(I_0=\varnothing\).  The diagonal convention
\(V_{ii}=\bar\sigma\) keeps only the disconnected vacuum-expectation
contribution.  The omitted diagonal connected finite parts are singular
crossed limits and require a separate diagonal finite-volume prescription.
\end{definition}

\begin{remark}[Connected TFFSA block]
\label{rem:zero-momentum-connected-tffsa-hermitian}
The matrix in Definition~\ref{def:zero-momentum-connected-tffsa-block} is
Hermitian.  If \(h=0\), its eigenvalues are
\[
  0,\quad 2\omega_0,\quad 2\omega_1,\quad\ldots,\quad
  2\omega_{N_{\rm pair}-1}.
\]
The diagonal entries \(E_i^{(0)}+hL\bar\sigma\) are real.  For \(i<j\) the
lower-triangular entry is defined to be \(\overline{V_{ij}}\), hence
\[
  H_{ji}=hL\,\overline{V_{ij}}=\overline{hL\,V_{ij}}=\overline{H_{ij}}
\]
for real \(h\) and \(L\).  Thus \(H^\dagger=H\).  At \(h=0\), all
off-diagonal entries vanish and the matrix is diagonal with the displayed
free finite-volume energies.
\end{remark}

The script
\[
  \texttt{qft\_scripts/tffsa\_ising\_spin\_connected.py}
\]
implements Definition~\ref{def:zero-momentum-connected-tffsa-block}.  It
checks Hermiticity in smoke mode and prints the finite eigenvalues together
with the corresponding free energies.  The purpose is to make the
form-factor normalization, density factors, and diagonal convention
executable.  It is not a claim that the small connected block is the full
magnetic-Ising spectrum.

\begin{lemma}[Finite TFFSA spectral-flow derivative]
\label{lem:finite-tffsa-hellmann-feynman}
Let
\begin{equation}
  H(h)=H_0+hW,\qquad H_0=H_0^\dagger,\quad W=W^\dagger ,
  \label{eq:finite-affine-hermitian-flow}
\end{equation}
be a finite Hermitian matrix family, with \(h\in\mathbb R\).  Suppose
\(\lambda_*(h_*)\) is a simple eigenvalue of \(H(h_*)\), and let \(u_*\) be a
normalized eigenvector.  If
\begin{equation}
  g=\operatorname{dist}\bigl(\lambda_*(h_*),
      \operatorname{spec}H(h_*)\setminus\{\lambda_*(h_*)\}\bigr)>0 ,
  \label{eq:finite-spectral-flow-gap}
\end{equation}
then for \(|h-h_*|\|W\|<g/2\) there is a unique eigenvalue branch
\(\lambda_*(h)\) near \(\lambda_*(h_*)\), counted with its normalized
eigenvector \(u(h)\) after fixing a differentiable phase.  Its derivative at
\(h=h_*\) is
\begin{equation}
  \frac{\dd\lambda_*}{\dd h}(h_*)
  =
  u_*^\dagger W u_* .
  \label{eq:finite-hellmann-feynman-derivative}
\end{equation}
If all eigenvalues of \(H(h_*)\) are simple and
\(\lambda_a'(h_*)=u_a^\dagger W u_a\), then
\begin{equation}
  \sum_a\lambda_a'(h_*)=\operatorname{tr} W .
  \label{eq:finite-spectral-flow-trace-identity}
\end{equation}
\end{lemma}

\begin{proof}
Choose a positively oriented circle \(C\) centered at \(\lambda_*(h_*)\) with
radius smaller than \(g/2\).  For \(z\in C\),
\[
  \|(h-h_*)W(H(h_*)-z)^{-1}\|
  \leq
  |h-h_*|\|W\|\,\|(H(h_*)-z)^{-1}\|
  <
  1 ,
\]
when \(|h-h_*|\|W\|<g/2\).  Thus the resolvent identity and the Neumann series
give a well-defined resolvent \((H(h)-z)^{-1}\) on \(C\), analytic in \(h\).
The Riesz projection
\[
  P(h)=\frac{1}{2\pi\ii}\oint_C (z-H(h))^{-1}\,\dd z
\]
therefore has rank one for \(h\) in this neighborhood and depends
differentiably on \(h\).  Taking \(u(h)=P(h)u_*/\|P(h)u_*\|\), with the phase
chosen so that \(u_*^\dagger u(h)\) is positive real near \(h_*\), gives a
differentiable normalized eigenvector for the unique eigenvalue branch inside
\(C\).

Differentiate
\[
  H(h)u(h)=\lambda_*(h)u(h)
\]
at \(h=h_*\).  Since \(H'(h)=W\),
\[
  Wu_*+H(h_*)u'(h_*)
  =
  \lambda_*'(h_*)u_*+\lambda_*(h_*)u'(h_*) .
\]
Taking the inner product with \(u_*\), and using
\(u_*^\dagger H(h_*)=\lambda_*(h_*)u_*^\dagger\), cancels the terms
containing \(u'(h_*)\) and gives
\eqref{eq:finite-hellmann-feynman-derivative}.  If all eigenvalues are simple,
their normalized eigenvectors form an orthonormal basis, so
\[
  \sum_a\lambda_a'(h_*)
  =
  \sum_a u_a^\dagger W u_a
  =
  \operatorname{tr}W .
\]
\end{proof}

For the connected Ising TFFSA block,
\eqref{eq:finite-affine-hermitian-flow} is exactly the finite matrix
\[
  H(h)=H_0+hL\,V .
\]
The companion script
\[
  \texttt{qft\_scripts/tffsa\_ising\_spectral\_flow.py}
\]
diagonalizes this finite block on a small grid of \(h\)-values and checks
\eqref{eq:finite-hellmann-feynman-derivative} against centered finite
differences at a point where the finite spectrum is separated.  This is a
useful nonintegrable benchmark because it checks the actual finite spectral
flow generated by spin-field form factors.  It remains a finite-matrix
check: comparison with the continuum magnetic-Ising spectrum requires
the cutoff, finite-volume, diagonal-form-factor, and counterterm estimates
described around \eqref{eq:tffsa-connected-spin-matrix-element}.

\subsection{\texorpdfstring{Exact \(E_8\) Target Data For Magnetic-Ising Benchmarks}{Exact E8 Target Data For Magnetic-Ising Benchmarks}}
\label{subsec:e8-ising-target-data}

The magnetic Ising ray \(\tau=0\), \(h\ne0\), is the standard case in which
Hamiltonian truncation is compared with exact integrable-QFT target data.
The target data are not obtained by diagonalizing a finite truncation.  They
come from the continuum factorized-scattering description of the integrable
massive theory.  In the numerical chapter we use only the part of that
description needed for a reproducible benchmark: the dimensionless mass
ratios
\[
  \rho_a=\frac{m_a}{m_1},\qquad a=1,\ldots,8,
\]
where \(m_1\) sets the overall mass scale determined by the normalization of
the magnetic coupling \(h\).  The eight ratios are
\begin{align}
  \rho_1 &= 1, &
  \rho_2 &= 2\cos\frac{\pi}{5}, &
  \rho_3 &= 2\cos\frac{\pi}{30}, \notag\\
  \rho_4 &= 2\rho_2\cos\frac{7\pi}{30}, &
  \rho_5 &= 2\rho_2\cos\frac{2\pi}{15}, &
  \rho_6 &= 2\rho_2\cos\frac{\pi}{30}, \notag\\
  \rho_7 &= 4\rho_2\cos\frac{\pi}{5}\cos\frac{7\pi}{30},&
  \rho_8 &= 4\rho_2\cos\frac{\pi}{5}\cos\frac{2\pi}{15}.
  \label{eq:e8-magnetic-ising-mass-ratios}
\end{align}
Numerically this is approximately
\begin{gather*}
  1,\quad
  1.6180339887,\quad
  1.9890437907,\quad
  2.4048671724,\\
  2.9562952015,\quad
  3.2183404585,\quad
  3.8911568233,\quad
  4.7833861168 .
\end{gather*}

There is a finite algebraic check for the convention used in
\eqref{eq:e8-magnetic-ising-mass-ratios}.  Let \(A_{E_8}\) be the adjacency
matrix of the \(E_8\) Dynkin graph with edges
\[
  (0,1),(1,2),(2,3),(3,4),(4,5),(5,6),(2,7).
\]
The largest eigenvalue is
\[
  \lambda_{\max}(A_{E_8})=2\cos\frac{\pi}{30}.
\]
If \(v\) is the positive Perron-Frobenius eigenvector and is normalized by
\(\min_i v_i=1\), then the unordered component ratios of \(v\) are precisely
the eight numbers in \eqref{eq:e8-magnetic-ising-mass-ratios}.  For the node
labeling above the normalized vector is
\[
  (v_0,\ldots,v_7)
  =
  (\rho_2,\rho_6,\rho_8,\rho_7,\rho_5,\rho_3,\rho_1,\rho_4).
\]
The eigenvector equations \(A_{E_8}v=\lambda_{\max}v\) reduce the statement
to elementary trigonometric identities such as
\[
  v_5+v_7=\lambda_{\max}v_6,\qquad
  v_1+v_3+v_7=\lambda_{\max}v_2,
\]
after substituting \(\lambda_{\max}=2\cos(\pi/30)\) and the ratios above.
This check fixes the table and its ordering; it does not replace the
continuum integrable-QFT derivation that the magnetic-Ising scattering theory
has this \(E_8\) particle content.

The companion script
\[
  \texttt{qft\_scripts/e8\_ising\_mass\_ratios.py}
\]
prints \eqref{eq:e8-magnetic-ising-mass-ratios} and verifies the
Perron--Frobenius check.  In a TCSA or TFFSA comparison the observable
being compared should be a set of finite-volume energy gaps
\(\Delta E_a(L,E_{\rm cut})\), divided by the lightest extracted gap, and
then extrapolated in \(L\) and \(E_{\rm cut}\) with a declared systematic
model.  Agreement with \eqref{eq:e8-magnetic-ising-mass-ratios} is a
target-data comparison only after those regulator coordinates have been
specified.

\begin{lemma}[Feshbach--Schur reduction with resolvent control]
\label{lem:hamiltonian-truncation-schur-complement}
In a finite regulated Hilbert space, let \(H=H^\dagger\) be a Hermitian
matrix and let \(\Hilb=P\Hilb\oplus Q\Hilb\), where \(P\) is an orthogonal
projection and \(Q=1-P\).  Suppose \(E\) is not an eigenvalue of the
finite matrix \(QHQ:Q\Hilb\to Q\Hilb\).  Then \(E\) is an eigenvalue of
\(H\) with \(P\psi\ne0\) only if
\begin{equation}
  \left[
    PHP
    -
    PHQ(QHQ-E)^{-1}QHP
  \right]P\psi
  =
  E\,P\psi .
  \label{eq:feshbach-schur-truncation-effective-hamiltonian}
\end{equation}
Conversely, if \(u\in P\Hilb\) satisfies
\eqref{eq:feshbach-schur-truncation-effective-hamiltonian}, then
\[
  \psi
  =
  u
  -
  (QHQ-E)^{-1}QHPu
\]
is an eigenvector of \(H\) with eigenvalue \(E\).
Moreover, if
\[
  d(E)=\operatorname{dist}(E,\operatorname{spec}(QHQ))>0,
\]
then the eliminated component obeys the finite-cutoff estimate
\begin{equation}
  \|Q\psi\|
  \leq
  d(E)^{-1}\,\|QHP\,P\psi\|.
  \label{eq:feshbach-q-component-bound}
\end{equation}
\end{lemma}

\begin{proof}
Write the eigenvalue equation \(H\psi=E\psi\) as two block equations:
\[
  (PHP-E)P\psi+PHQ\,Q\psi=0,
  \qquad
  QHP\,P\psi+(QHQ-E)Q\psi=0 .
\]
Since \(E\) is in the resolvent set of \(QHQ\), the second equation gives
\[
  Q\psi=-(QHQ-E)^{-1}QHP\,P\psi .
\]
Substituting this into the first equation gives
\eqref{eq:feshbach-schur-truncation-effective-hamiltonian}.  This proves that
any eigenvector with nonzero \(P\)-projection gives a vector in the kernel of
the energy-dependent effective operator on \(P\Hilb\).  Conversely, suppose
\(u\in P\Hilb\) obeys the Schur-complement equation and define
\[
  \psi=u-(QHQ-E)^{-1}QHPu .
\]
Then \(P\psi=u\) and the second block equation is satisfied by construction,
because \((QHQ-E)(QHQ-E)^{-1}=Q\) on \(Q\Hilb\).  Substituting the
Schur-complement equation into the first block equation gives zero there as
well.  Hence \(H\psi=E\psi\).

Since \(QHQ\) is a finite Hermitian matrix, the spectral theorem gives
\[
  \|(QHQ-E)^{-1}\|=d(E)^{-1}.
\]
Applying this to the explicit formula for \(Q\psi\) gives
\eqref{eq:feshbach-q-component-bound}.  This estimate is the part of the
finite-dimensional identity that survives as useful information in truncation
theory: a high-energy block can be eliminated only together with control of
the resolvent distance and the coupling \(QHP\).
\end{proof}

Equation~\eqref{eq:feshbach-schur-truncation-effective-hamiltonian} is the
precise operator identity behind cutoff counterterms in TCSA and TFFSA.  In
practice one expands the \(Q\)-space resolvent at high energy and replaces its
low-energy action by local or quasilocal operators allowed by the symmetries.
For the Ising spin perturbation this expansion is governed by the short-
distance operator product of \(\sigma\) with itself; for the free-fermion
basis it is equivalently encoded in the high-rapidity behavior of the spin
form factors.  A claimed numerical spectrum is therefore meaningful only
after the truncation scheme states the cutoff, the finite-volume form-factor
prescription, the symmetry sector, the counterterm coordinates, and the
extrapolation model.

\section{Finite-Matrix Spectral Bounds}
\label{sec:finite-matrix-spectral-bounds}

Every Hamiltonian-truncation or DLCQ computation ultimately diagonalizes a
finite Hermitian matrix.  This elementary fact is useful only if the finite
linear-algebra output is verified separately from the continuum claim.  We
therefore record the minimal residual bound used by the companion scripts.

\begin{definition}[Residual bound for a finite truncation]
\label{def:finite-truncation-residual-bound}
Let \(A=A^\dagger\) be a finite Hermitian matrix on a finite-dimensional
Hilbert space \(V\).  A residual bound for a proposed eigenpair is a
triple
\begin{equation}
  (\lambda,u,\rho),\qquad \|u\|=1,\qquad
  \rho=\|(A-\lambda)u\|.
  \label{eq:finite-matrix-residual-bound}
\end{equation}
It is a bound for the finite matrix \(A\) only.  The number
\(\lambda\) becomes a QFT energy or invariant mass only after the regulator
label, counterterm prescription, symmetry sector, finite-volume convention,
and continuum-extrapolation hypothesis have also been supplied.
\end{definition}

\paragraph{Spectral meaning of a finite residual.}
Let \(A=A^\dagger\) be a finite Hermitian matrix, let
\(\|u\|=1\), and set \(\rho=\|(A-\lambda)u\|\).
Then
\begin{equation}
  \operatorname{dist}(\lambda,\operatorname{spec}A)\leq \rho .
  \label{eq:residual-distance-to-spectrum}
\end{equation}
More generally, let \(S\subset\operatorname{spec}A\) and let \(P_S\) be the
orthogonal spectral projection onto the span of eigenvectors whose
eigenvalues lie in \(S\).  If
\begin{equation}
  g=\operatorname{dist}\bigl(\lambda,\operatorname{spec}A\setminus S\bigr)>0,
  \label{eq:finite-spectral-gap-from-complement}
\end{equation}
then
\begin{equation}
  \|(1-P_S)u\|\leq \frac{\rho}{g}.
  \label{eq:finite-projector-leakage-bound}
\end{equation}

Choose an orthonormal eigenbasis \(e_a\) of \(A\),
\[
  Ae_a=E_a e_a,\qquad u=\sum_a c_a e_a,\qquad \sum_a |c_a|^2=1.
\]
Then
\begin{equation}
  \rho^2
  =
  \|(A-\lambda)u\|^2
  =
  \sum_a |c_a|^2 |E_a-\lambda|^2 .
  \label{eq:residual-spectral-expansion}
\end{equation}
If every eigenvalue obeyed \(|E_a-\lambda|>\rho\), the right hand side of
\eqref{eq:residual-spectral-expansion} would be strictly larger than
\(\rho^2\), a contradiction.  This proves
\eqref{eq:residual-distance-to-spectrum}.  For the projection bound, split
the sum into \(E_a\in S\) and \(E_a\notin S\).  On the complement,
\(|E_a-\lambda|\geq g\), hence
\[
  \rho^2
  \geq
  g^2\sum_{E_a\notin S}|c_a|^2
  =
  g^2\|(1-P_S)u\|^2 .
\]
Taking square roots gives
\eqref{eq:finite-projector-leakage-bound}.

The first inequality says that a small residual forces the existence of some
finite-matrix eigenvalue near \(\lambda\).  The second says that, when a
target spectral cluster is separated from the rest of the finite spectrum,
the approximate eigenvector has quantitatively small leakage out of that
cluster.  These statements do not estimate the distance between different
cutoffs.  Cutoff comparison belongs to the Feshbach identity above, to
counterterm matching, and to the extrapolation bounds below.

For a Hermitian block decomposition \(V=PV\oplus QV\), the same finite
linear algebra gives a practical diagnostic for a truncation counterterm.
At a chosen low-energy value \(E\), the omitted-space correction is exactly
\begin{equation}
  \Sigma_P(E)=PHQ(QHQ-E)^{-1}QHP .
  \label{eq:finite-schur-self-energy}
\end{equation}
Approximating \(\Sigma_P(E)\) by a low-dimensional local counterterm is a
separate approximation; the identity itself is exact only before this
replacement is made.  A controlled truncation must therefore state a norm,
an energy window, and a bound on
\begin{equation}
  \Sigma_P(E)-\Sigma_P^{\rm ct}(E)
  \label{eq:schur-counterterm-remainder}
\end{equation}
in that window.

The companion check
\[
  \texttt{calculation-checks/hamiltonian\_truncation\_dlcq\_checks.py}
\]
verifies the finite residual bound
\eqref{eq:residual-distance-to-spectrum}, the projector leakage estimate
\eqref{eq:finite-projector-leakage-bound}, the Feshbach determinant identity
behind \eqref{eq:feshbach-schur-truncation-effective-hamiltonian}, the
Ising-energy Bogoliubov benchmark, the connected Ising TFFSA block
normalization, the finite Ising TFFSA spectral-flow derivative identity, the
\(E_8\) magnetic-Ising target-ratio Perron--Frobenius check, the finite
large-\(N\) QCD DLCQ quadratic-form identity, and the exact-intercept cases
of the finite large-\(K\) extrapolation diagnostic.

\section{Krylov Data as Finite Spectral Evidence}
\label{sec:krylov-finite-spectral-evidence}

Large truncation matrices are rarely diagonalized by forming a complete
eigenbasis.  The usual numerical output is a Krylov subspace calculation:
Lanczos vectors, a tridiagonal compression, residuals, and sometimes a
spectral density with respect to a chosen seed vector.  These objects have
exact finite-matrix meanings.  They become evidence for a QFT spectrum only
after the finite regulator, symmetry sector, seed overlap, counterterms, and
continuum extrapolation are separately controlled.

\begin{definition}[Krylov spectral data]
\label{def:krylov-spectral-data}
Let \(A=A^\dagger\) be a finite Hermitian matrix on a finite-dimensional
Hilbert space \(V\), and let \(v\in V\) be a normalized seed vector.  For
\(m\geq1\), the \(m\)-th Krylov subspace is
\[
  \mathcal K_m(A,v)
  =
  \operatorname{span}\{v,Av,\ldots,A^{m-1}v\}.
\]
A Krylov spectral datum consists of an orthonormal basis
\[
  V_m=(q_0,\ldots,q_{m-1})
  \quad\text{of}\quad
  \mathcal K_m(A,v),
  \qquad q_0=v,
\]
the compression
\[
  T_m=V_m^\dagger A V_m ,
\]
and the residual matrix
\[
  R_m=A V_m-V_mT_m .
\]
When the basis is the exact Lanczos basis before breakdown, \(T_m\) is
tridiagonal and
\begin{equation}
  A V_m=V_mT_m+\beta_m q_m e_m^\dagger ,
  \label{eq:lanczos-residual-relation}
\end{equation}
where \(q_m\perp\mathcal K_m(A,v)\), \(\|q_m\|=1\), \(e_m\) is the last
coordinate vector in \(\mathbb C^m\), and \(\beta_m\geq0\).  Breakdown
means \(\beta_m=0\), in which case \(\mathcal K_m(A,v)\) is already an
invariant subspace for \(A\).
\end{definition}

Suppose \(T_m y=\theta y\), \(\|y\|=1\), and set \(u=V_m y\).  Multiplying
\eqref{eq:lanczos-residual-relation} by \(y\) gives the exact finite
residual identity
\begin{equation}
  (A-\theta)u=\beta_m\, y_m\, q_m,
  \qquad
  \|(A-\theta)u\|=\beta_m |y_m| .
  \label{eq:lanczos-ritz-residual-bound}
\end{equation}
Thus a reported Lanczos Ritz value is accompanied by a finite-matrix
residual bound of the kind defined in
Definition~\ref{def:finite-truncation-residual-bound}.  Its content is
finite and sectorwise: the regulated matrix in the declared sector has
spectrum within \(\beta_m|y_m|\) of \(\theta\), and the approximate
eigenvector leakage into a separated finite spectral cluster is bounded by
\eqref{eq:finite-projector-leakage-bound}.  A continuum QFT spectral claim
then requires the regulator and extrapolation data specified in
Definition~\ref{def:hamiltonian-truncation-regulator-datum}.

The same Krylov data carry finite spectral-measure information.  Let
\[
  \mu_v=\sum_a |\langle e_a,v\rangle|^2\delta_{E_a}
\]
be the finite spectral measure of \(A\) in the seed state, where
\(\{e_a\}\) is an orthonormal eigenbasis of \(A\), \(Ae_a=E_a e_a\).  If
\(\theta_r\) are the eigenvalues of \(T_m\) and \(s^{(r)}\) are normalized
eigenvectors of \(T_m\), define
\[
  w_r=|\langle e_1,s^{(r)}\rangle|^2 ,
  \qquad
  \mu_{v,m}^{\rm Krylov}=\sum_{r=1}^{m}w_r\delta_{\theta_r}.
\]
For every polynomial \(p\) of degree at most \(2m-1\),
\begin{equation}
  \int p(\lambda)\,\dd\mu_v(\lambda)
  =
  \sum_{r=1}^{m} w_r p(\theta_r).
  \label{eq:krylov-gauss-moment-exactness}
\end{equation}
To see the mechanism, first note that if \(\deg b\leq m-1\), then
\[
  b(A)v=V_m b(T_m)e_1 .
\]
Let \(\pi_m(\lambda)=\det(\lambda I-T_m)\).  The Lanczos recurrence implies
that \(\pi_m(A)v\) is orthogonal to \(\mathcal K_m(A,v)\), while
\(\pi_m(T_m)=0\).  Any polynomial \(p\) with \(\deg p\leq2m-1\) decomposes
as \(p=a\pi_m+b\) with \(\deg a,\deg b\leq m-1\).  Since
\(a(A)v\in\mathcal K_m(A,v)\),
\[
  \langle v,p(A)v\rangle
  =
  \langle v,b(A)v\rangle
  =
  e_1^\dagger b(T_m)e_1
  =
  e_1^\dagger p(T_m)e_1,
\]
and diagonalizing \(T_m\) gives
\eqref{eq:krylov-gauss-moment-exactness}.  The finite Krylov measure is
therefore an exact quadrature rule for the seed spectral measure, not a
complete list of all finite eigenvalues unless the chosen seeds have overlap
with the relevant spectral subspaces.

Symmetry sectors must be part of the same finite spectral datum.  If a finite group or
compact group acts by unitary matrices commuting with \(A\), and \(P_\alpha\)
is the exact projector onto a declared irreducible sector, then the Krylov
calculation for that sector is the calculation for the finite matrix
\[
  A_\alpha=P_\alpha A P_\alpha
  \quad\text{on}\quad
  P_\alpha V .
\]
A seed \(v\) with \(P_\alpha v=0\) contains no spectral information about
that sector, and a single nonzero seed can miss degenerate eigenvectors
orthogonal to its cyclic subspace.  A numerical claim about absence of states
in a finite window must therefore use either full diagonalization in the
sector, a block-Krylov seed family with a rank check for the target
subspace, or an independent variational lower bound.  Without such sector
coverage, Lanczos data are reliable spectral evidence for the cyclic
subspace generated by the stated seed, not for the whole Hilbert space.

\section{Variational Ansatz Bounds}
\label{sec:variational-ansatz-bounds}

Tensor-network, DMRG, MERA, and neural-state computations enter QFT as
finite-regulator variational methods.  The ansatz may be sophisticated, but
the first mathematical object is still a finite Hilbert space, a finite
Hermitian operator, and a family of trial vectors.  The following finite
bounds are the common core behind those methods.

\begin{definition}[Finite variational ansatz]
\label{def:finite-variational-ansatz-datum}
A finite variational ansatz for a regulated Hamiltonian consists of
\[
  (V,A,\Theta,\Psi,\mathcal S,\mathcal E_{\rm samp}) .
\]
Here \(V\) is a finite-dimensional Hilbert space, \(A=A^\dagger\) is the
finite Hamiltonian or invariant-mass matrix in a declared sector,
\(\Theta\) is the parameter space, and
\[
  \Psi:\Theta\to V\setminus\{0\}
\]
is the trial-vector map.  The entry \(\mathcal S\) records exact constraints
imposed on the ansatz, such as gauge invariance, global charge, momentum,
parity, tensor-network bond dimension, locality of tensor contractions, or a
neural architecture.  The entry \(\mathcal E_{\rm samp}\) records the
sampling law and estimator if the Rayleigh quotient is estimated
stochastically rather than contracted exactly.
\end{definition}

For \(\psi\neq0\), write
\[
  E_\psi=\frac{\langle\psi,A\psi\rangle}{\langle\psi,\psi\rangle},
  \qquad
  r_\psi=(A-E_\psi)\psi,
  \qquad
  \Delta_\psi^2=\frac{\|r_\psi\|^2}{\|\psi\|^2}.
\]
The number \(\Delta_\psi^2\) is the energy variance in the trial state:
\[
  \Delta_\psi^2
  =
  \frac{\langle\psi,A^2\psi\rangle}{\langle\psi,\psi\rangle}
  -
  E_\psi^2 .
\]
Consequently
\begin{equation}
  \operatorname{dist}(E_\psi,\operatorname{spec}A)\leq \Delta_\psi .
  \label{eq:variational-variance-spectrum-distance}
\end{equation}
This is exactly the residual bound
\eqref{eq:residual-distance-to-spectrum} applied to
\(\psi/\|\psi\|\).  If \(S\subset\operatorname{spec}A\) is a separated
finite spectral cluster and
\[
  g=\operatorname{dist}(E_\psi,\operatorname{spec}A\setminus S)>0 ,
\]
then
\begin{equation}
  \|(1-P_S)\psi\|
  \leq
  \frac{\Delta_\psi}{g}\,\|\psi\| .
  \label{eq:variational-variance-projector-leakage}
\end{equation}
Thus variance is a finite-matrix eigenstate diagnostic.  It measures
spectral concentration of the vector already produced by the ansatz; it does
not by itself measure the expressive power of the ansatz family.

The Rayleigh quotient supplies a different bound when an exact
lower spectral edge is available.  Let
\[
  E_0<E_1
\]
be the lowest two distinct eigenvalues of \(A\), and let \(P_0\) be the
ground-space projection.  Expanding a normalized vector in an eigenbasis gives
\[
  E_\psi-E_0
  =
  \sum_a |\langle e_a,\psi\rangle|^2(E_a-E_0)
  \geq
  (E_1-E_0)\|(1-P_0)\psi\|^2 .
\]
Therefore
\begin{equation}
  \|(1-P_0)\psi\|^2
  \leq
  \frac{E_\psi-E_0}{E_1-E_0}.
  \label{eq:variational-ground-projector-bound}
\end{equation}
This bound is useful only after \(E_0\) and a finite gap are known by another
method, or after a rigorous lower bound replaces the exact ground energy.
The variational upper bound \(E_\psi\geq E_0\) alone is insufficient to
bound distance to the ground-state subspace.

For a smooth normalized ansatz \(\psi(\theta)\), the differential of the
Rayleigh quotient exposes the precise meaning of a variational optimum.  If
\(\|\psi(\theta)\|=1\) and \(\xi\in T_\theta\Theta\), put
\(\dot\psi_\xi=d\psi_\theta(\xi)\).  Differentiating
\(\langle\psi,A\psi\rangle\) and using
\(\operatorname{Re}\langle\dot\psi_\xi,\psi\rangle=0\) gives
\begin{equation}
  dE_\theta(\xi)
  =
  2\,\operatorname{Re}
  \langle \dot\psi_\xi,(A-E_\theta)\psi(\theta)\rangle .
  \label{eq:variational-tangent-gradient}
\end{equation}
A stationary variational parameter therefore makes the residual orthogonal
to the tangent space of the ansatz manifold.  An eigenvector error bound is
stronger: it requires the residual norm \(\|(A-E_\theta)\psi(\theta)\|\) to
be small, or a separate argument that the tangent directions cover the
missing finite directions.

For neural-state and Monte Carlo variational calculations one also needs the
finite estimator identity.  Choose an orthonormal computational basis
\(\{|x\rangle\}_{x\in X}\) and assume \(\psi_x=\langle x|\psi\rangle\neq0\)
on the sampled support.  Define
\[
  p_\psi(x)=\frac{|\psi_x|^2}{\sum_y|\psi_y|^2},
  \qquad
  E_{\rm loc}(x)=\frac{(A\psi)_x}{\psi_x}.
\]
Then
\begin{equation}
  E_\psi=\sum_x p_\psi(x)E_{\rm loc}(x),
  \label{eq:variational-local-energy-mean}
\end{equation}
and
\begin{equation}
  \Delta_\psi^2
  =
  \sum_x p_\psi(x)|E_{\rm loc}(x)-E_\psi|^2 .
  \label{eq:variational-local-energy-variance}
\end{equation}
The second identity uses Hermiticity of \(A\):
\[
  \sum_x p_\psi(x)|E_{\rm loc}(x)-E_\psi|^2
  =
  \frac{\|(A-E_\psi)\psi\|^2}{\|\psi\|^2}.
\]
Sampling estimates of \eqref{eq:variational-local-energy-mean} and
\eqref{eq:variational-local-energy-variance} require the Markov-chain or
independent-sample error analysis of
Section~\ref{sec:blocking-jackknife-bootstrap-error-coordinates}; optimization
bias, autocorrelation, and architecture bias are additional entries of
\(\mathcal E_{\rm samp}\), not properties of the finite identities.

The force used in a smooth neural-state or variational-Monte-Carlo
optimization is an equally finite identity.  Let
\(\theta=(\theta^a)\) be real local coordinates on a patch of \(\Theta\), and
assume \(\psi_x(\theta)\neq0\) on the sampled support.  Define the score
coordinate
\[
  \mathsf s_a(x;\theta)=\partial_a\log\psi_x(\theta).
\]
Put \(p_\theta=p_{\psi(\theta)}\) and
\(E_{\rm loc}(x;\theta)=((A\psi(\theta))_x)/\psi_x(\theta)\).
Without normalizing the vector \(\psi(\theta)\), differentiation of the
Rayleigh quotient gives
\[
  \partial_a E_\theta
  =
  2\,\operatorname{Re}
  \frac{\langle \partial_a\psi,(A-E_\theta)\psi\rangle}
       {\langle\psi,\psi\rangle}.
\]
In the computational basis this becomes the score--local-energy covariance
identity
\begin{equation}
  \partial_a E_\theta
  =
  2\,\operatorname{Re}
  \sum_x p_\theta(x)
  \Bigl(
    \overline{\mathsf s_a(x;\theta)}
    -
    \sum_y p_\theta(y)\overline{\mathsf s_a(y;\theta)}
  \Bigr)
  \bigl(E_{\rm loc}(x;\theta)-E_\theta\bigr).
  \label{eq:variational-score-force-identity}
\end{equation}
The centering of the score does not change the value because
\(\sum_xp_\theta(x)(E_{\rm loc}(x;\theta)-E_\theta)=0\).  Equation
\eqref{eq:variational-score-force-identity} is what a stochastic force
estimator estimates after a sampling law has been declared.  It is not an
optimizer theorem: a small empirical force may indicate a tangent-space
stationary point while the residual norm, the energy variance, the integrated
autocorrelation time of the Markov chain, or the ansatz bias remains large.
For QFT applications the finite force is therefore reported together with
\eqref{eq:variational-local-energy-variance}, sector constraints, and the
regulator extrapolation data of
Section~\ref{sec:finite-regulator-extrapolation}.

Tensor-network terminology fits this definition by specifying \(\Psi\).  A
finite matrix-product state on a spin chain of length \(L\), physical local
dimension \(d\), and bond dimension \(\chi\) is a vector whose amplitudes
are finite contractions of matrices
\[
  \psi_{i_1\ldots i_L}
  =
  \operatorname{Tr}\!\left(
    A^{i_1}_1A^{i_2}_2\cdots A^{i_L}_L
  \right),
  \qquad
  A^i_\ell\in\operatorname{Mat}_{\chi_{\ell-1}\times\chi_\ell}(\mathbb C).
\]
For an infinite translation-invariant MPS, a transfer-operator spectrum can
estimate a correlation length, but a QFT continuum claim still requires a
sequence of finite regulators, a scaling map for operators, and convergence
of the corresponding Schwinger or Wightman data.  Bond dimension, discarded
weight, transfer-operator gap, local-energy variance, and Krylov residuals
are finite evidence coordinates.  The continuum claim is the additional
statement that those coordinates control the regulator limit in a declared
topology.

\paragraph{Finite transfer-operator correlation length.}
The most common tensor-network continuum diagnostic is the transfer-operator
gap.  Here is the finite statement in the form needed for QFT use.  For a
translation-invariant infinite MPS with tensors \(A^i\), define the finite
transfer operator
\[
  \mathbb E(X)
  =
  \sum_i A^i X (A^i)^\dagger
\]
on \(\operatorname{Mat}_{\chi\times\chi}(\mathbb C)\).  Assume, for this
finite ansatz, that \(\mathbb E\) has a simple eigenvalue
\(\lambda_0=1\), positive left and right eigenvectors \(\ell,r\) normalized
by \(\ell(r)=1\), and remaining eigenvalues
\(\lambda_\alpha\) with \(|\lambda_\alpha|<1\).  A local lattice operator
\(O\) is represented by the insertion map
\[
  \mathbb E_O(X)
  =
  \sum_{i,j}O_{ij}A^iX(A^j)^\dagger .
\]
The connected two-point function at lattice separation \(n\ge1\) is
\begin{equation}
  C_O(n)
  =
  \ell\!\left(\mathbb E_O\mathbb E^{\,n}\mathbb E_O r\right)
  -
  \ell(\mathbb E_Or)^2 .
  \label{eq:mps-transfer-connected-correlator}
\end{equation}
If \(\mathbb E\) is diagonalizable on the subspace complementary to the
fixed vector, then
\begin{equation}
  C_O(n)
  =
  \sum_\alpha A_\alpha(O)\lambda_\alpha^{\,n},
  \label{eq:mps-transfer-spectral-expansion}
\end{equation}
where the coefficients are products of the left and right spectral
projections of \(\mathbb E_O\).  Thus a single isolated contributing
eigenvalue with \(\lambda_\alpha>0\) gives a finite lattice correlation
length
\[
  \xi_\alpha^{\rm lat}
  =
  -\frac1{\log\lambda_\alpha}.
\]
If the lattice spacing is \(a\), the corresponding finite effective mass is
\begin{equation}
  m_\alpha(a)
  =
  -\frac1a\log\lambda_\alpha .
  \label{eq:mps-transfer-effective-mass}
\end{equation}

This is a finite transfer-operator identity, not yet a continuum QFT
statement.  To interpret \(m_\alpha(a)\) as a physical mass one needs a
sequence of regulator data \((a,\chi(a),A^i(a),O_a)\), a normalization of the
operator \(O_a\), and estimates of the form
\begin{equation}
  \lambda_\alpha(a)
  =
  \exp[-a m+O(a^{1+\delta})],
  \qquad
  Z_O(a)^2 A_\alpha(O_a)\longrightarrow A_{\rm phys},
  \label{eq:mps-transfer-continuum-input}
\end{equation}
together with convergence of the corresponding multipoint Schwinger or
Wightman distributions.  Excited transfer modes are not nuisances in this
language; they are visible finite terms in
\eqref{eq:mps-transfer-spectral-expansion}.  If
\[
  C_O(n)=A_1e^{-a m_1n}+A_2e^{-a m_2n},
  \qquad m_2>m_1,\quad A_1,A_2>0,
\]
then the finite effective mass
\[
  m_{\rm eff}(a,n)
  =
  -\frac1a\log\frac{C_O(n+1)}{C_O(n)}
\]
lies between \(m_1\) and \(m_2\) and approaches \(m_1\) only after the
second transfer mode is suppressed.  A finite transfer gap therefore tells
one which exponential appears in the finite ansatz; it does not, by itself,
prove the continuum operator map, reflection positivity, locality, or
cutoff-removal estimate needed for QFT.

\paragraph{Calculation check.}
The companion check
\[
  \texttt{calculation-checks/hamiltonian\_truncation\_dlcq\_checks.py}
\]
also verifies the finite variational identities
\eqref{eq:variational-variance-spectrum-distance}--
\eqref{eq:variational-score-force-identity}: energy variance as residual
norm, spectral distance bounded by variance, ground-projection leakage from a
known finite gap, the tangent-gradient formula, and the local-energy
Monte Carlo and score-force identities in a finite basis.  It also checks the finite
transfer-operator spectral expansion
\eqref{eq:mps-transfer-spectral-expansion}, the single-mode effective-mass
identity \eqref{eq:mps-transfer-effective-mass}, the two-mode contamination
bound for \(m_{\rm eff}(a,n)\), and the elementary logarithmic remainder
bound implicit in \eqref{eq:mps-transfer-continuum-input}.

\section{Light-Front Kinematics And Constraints}
\label{sec:light-front-kinematic-constraint-datum}

Light-front Hamiltonian methods use a null hyperplane as the time slice.  In
the mostly-plus convention, choose one spatial direction \(x^1\) and set
\begin{equation}
  x^\pm=\frac{x^0\pm x^1}{\sqrt2},
  \qquad
  p^\pm=\frac{p^0\pm p^1}{\sqrt2}.
  \label{eq:general-light-front-coordinates}
\end{equation}
For \(D\)-dimensional Minkowski space,
\begin{equation}
  \dd s^2=-2\,\dd x^+\dd x^-+\dd x_\perp^2,
  \qquad
  p\cdot x=-p^-x^+-p^+x^-+p_\perp\cdot x_\perp,
  \label{eq:general-light-front-metric}
\end{equation}
and \(x^+\) is the evolution parameter.  The generator conjugate to \(x^+\)
is \(P^-\); the longitudinal momentum \(P^+\) and transverse momenta
\(P_\perp\) label superselection sectors for the finite light-front
Hamiltonians used below.  These identities are kinematic; they do not define
an interacting QFT until the constraint and zero-mode sectors have also been
specified.

\begin{definition}[Light-front Hamiltonian regulator specification]
\label{def:light-front-hamiltonian-regulator-datum}
A light-front Hamiltonian regulator specification is a Hamiltonian truncation
regulator specification
\(\mathfrak H_{\rm trunc}\) together with the following additional
structure.
\begin{enumerate}[label=\textnormal{(\roman*)}]
\item A null-coordinate chart \eqref{eq:general-light-front-coordinates},
  the sign convention \eqref{eq:general-light-front-metric}, and a declared
  evolution parameter \(x^+\).
\item A positive-longitudinal nonzero-mode sector, with one-particle labels
  carrying \(p^+>0\) and whatever transverse, spin, flavor, color, and
  statistics labels are kept by the regulator.
\item A zero-mode and constraint sector
  \[
    \mathcal Z_\Lambda,\qquad
    \mathcal C_{\Lambda,\alpha}=0,\qquad
    \mathcal G_{\Lambda,\rm res},
  \]
  recording the \(p^+=0\) modes, finite global Gauss-law constraints,
  residual gauge transformations, and any additional compact or topological
  degrees of freedom.  Eliminating, quantizing, or imposing these variables
  are different regulator choices and must be part of the regulator
  specification.
\item A finite physical Hilbert space
  \[
    \Hilb^{\rm phys}_\Lambda
    \subset
    \Hilb^{p^+>0}_\Lambda\otimes\Hilb_{\mathcal Z,\Lambda}
  \]
  obtained after the stated constraint prescription, together with commuting
  finite operators \(P^+_\Lambda\) and \(P_{\perp,\Lambda}\) on the physical
  sector.
\item A finite self-adjoint light-front Hamiltonian \(P^-_\Lambda\) on
  \(\Hilb^{\rm phys}_\Lambda\), commuting with the retained translation
  generators on the finite physical sector.  In a simultaneous
  \(P^+=p^+>0\), \(P_\perp=p_\perp\) sector, the finite invariant-mass
  operator is
  \begin{equation}
    M_\Lambda^2
    =
    2p^+P^-_\Lambda-p_\perp^2.
    \label{eq:light-front-mass-operator-fixed-sector}
  \end{equation}
  Outside a fixed momentum sector the same formula is read as a joint
  functional calculus expression on the common spectral representation of the
  commuting momenta and \(P^-_\Lambda\).
\item A continuum diagnostic specifying how \(p^+\)-cutoffs, harmonic
  resolution, transverse cutoffs, counterterms, principal-value
  prescriptions, and zero-mode constraints are taken to their limiting
  values.
\end{enumerate}
\end{definition}

\paragraph{One-particle measure.}
The free massive representation fixes the basic measure behind the
positive-\(p^+\) Fock construction.  For a test function \(F\) on momentum
space and \(m>0\),
\begin{equation}
  \int_{\mathbb R^D}\dd^Dp\,
  \theta(p^0)\delta(p^2+m^2)F(p)
  =
  \int_{p^+>0}
  \frac{\dd p^+\,\dd^{D-2}p_\perp}{2p^+}\,
  F\!\left(p^+,\,\frac{m^2+p_\perp^2}{2p^+},\,p_\perp\right).
  \label{eq:light-front-one-particle-measure}
\end{equation}
Indeed,
\[
  p^2+m^2=-2p^+p^-+p_\perp^2+m^2,
\]
so the \(p^-\)-integration localizes at
\begin{equation}
  p^-=\frac{m^2+p_\perp^2}{2p^+}
  \label{eq:light-front-mass-shell-relation}
\end{equation}
with Jacobian \(|\partial(p^2+m^2)/\partial p^-|=2p^+\).  The
positive-energy condition selects \(p^+>0\) on the massive shell.  The
massless boundary \(p^+=0\), and the constrained zero modes of gauge
theories, are not included in this one-particle measure; they are entries of
\(\mathcal Z_\Lambda\) and \(\mathcal C_{\Lambda,\alpha}\) in
Definition~\ref{def:light-front-hamiltonian-regulator-datum}.

\paragraph{Spectral interpretation.}
At fixed \(p^+\) and \(p_\perp\), \(M_\Lambda^2\) in
\eqref{eq:light-front-mass-operator-fixed-sector} and \(P^-_\Lambda\) have
the same finite eigenvectors, and their eigenvalues are related by
\[
  M^2=2p^+P^- -p_\perp^2.
\]
This is the finite-sector reason light-front calculations often diagonalize
\(P^-\) while reporting invariant masses.  A gauge-theory light-front
calculation is therefore specified only after the nonzero-mode Hamiltonian,
zero-mode/constraint prescription, and residual gauge data have all been
declared in the same regulator scheme.

\section{DLCQ Regulator}

DLCQ is the compact positive-longitudinal-momentum specialization of
Definition~\ref{def:light-front-hamiltonian-regulator-datum}.  It
compactifies
\[
  x^-\sim x^-+2\pi R
\]
so that
\[
  p^+=\frac{n}{R},\qquad n\in\mathbb Z_{\ge0}.
\]
The total longitudinal momentum is
\[
  P^+=\frac{K}{R},
\]
where \(K\) is the harmonic resolution.  The finite DLCQ Hilbert space at
fixed \(K\) consists of partitions of \(K\) among partons, with transverse and
internal cutoffs added as needed.  Zero modes with \(n=0\) are constrained
variables or additional degrees of freedom depending on the theory; their
treatment is part of the regulator.

\begin{definition}[DLCQ regulator specification]
\label{def:dlcq-regulator-datum}
A DLCQ regulator specification is a Hamiltonian truncation regulator
specification together
with the following light-front entries.
\begin{enumerate}[label=\textnormal{(\roman*)}]
\item The spatial light-front coordinate is compactified as
  \(x^-\sim x^-+2\pi R\), and the total longitudinal momentum sector is fixed
  by \(P^+=K/R\) with harmonic resolution \(K\in\mathbb Z_{>0}\).
\item The finite basis consists of all declared parton partitions
  \(K=n_1+\cdots+n_N\), with \(n_i>0\), together with the transverse,
  internal, gauge, and statistics labels kept by the regulator.
\item The zero-mode sector is specified as part of the regulator: constrained
  variables, eliminated modes, residual global constraints, and additional
  zero-mode degrees of freedom are all recorded before diagonalization.
\item The finite operator is the light-front Hamiltonian \(P^-_K\), or
  equivalently the invariant mass operator
  \(M_K^2=2P^+P^-_K-P_\perp^2\), with counterterms and principal-value
  prescriptions stated in the same regulator scheme.
\item The continuum diagnostic specifies the ordered finite labels followed as
  \(K\to\infty\), the extrapolation variable, and the remainder model used to
  interpret finite-\(K\) data.
\end{enumerate}
\end{definition}

\section{\texorpdfstring{Scalar \(\phi^4\) DLCQ Warmup}{Scalar phi4 DLCQ Warmup}}
\label{sec:scalar-phi4-dlcq-warmup}

Before the large-\(N\) gauge-theory example, it is useful to record a
nonintegrable scalar light-front matrix whose finite definition is completely
explicit.  This is a warmup for DLCQ, not a definition of the continuum
\(\phi^4_2\) theory.  The \(p^+=0\) mode is omitted in the construction
below; in light-front scalar theory that omission is a regulator choice and
is precisely where vacuum and symmetry-breaking information can hide.  A
continuum claim therefore needs a zero-mode prescription and a
counterterm/extrapolation theorem in addition to the finite matrix.

\begin{construction}[Finite scalar DLCQ matrix]
\label{con:finite-scalar-phi4-dlcq-matrix}
Fix a harmonic resolution \(K\in\mathbb Z_{>0}\), a mass \(m>0\), and a
finite coupling coordinate \(g\).  The finite bosonic basis consists of
occupation vectors
\begin{equation}
  {\bf N}=(N_1,\ldots,N_K)\in\mathbb Z_{\ge0}^K,
  \qquad
  \sum_{n=1}^K nN_n=K .
  \label{eq:phi4-dlcq-partition-basis}
\end{equation}
The parton with label \(n\) carries longitudinal momentum \(p_n^+=n/R\), and
the total sector has \(P^+=K/R\).  The free invariant mass is
\begin{equation}
  M^2_{0,K}({\bf N})
  =
  K m^2\sum_{n=1}^K\frac{N_n}{n}.
  \label{eq:phi4-dlcq-free-mass}
\end{equation}
This is the finite-sector form of
\[
  M^2=2P^+P^-,
  \qquad
  P^-_0=\sum_{n=1}^K\frac{m^2}{2p_n^+}\,N_n,
\]
with zero transverse momentum.

Let \(a_n,a_n^\dagger\) be bosonic creation and annihilation operators with
\([a_n,a_m^\dagger]=\delta_{nm}\).  The dimensionless normal-ordered quartic
operator \(Q_K\) is
\begin{equation}
  Q_K
  =
  \sum_{\substack{\epsilon_i=\pm1,\ 1\le n_i\le K\\
                  \sum_i\epsilon_i n_i=0}}
  \frac{
    :B_{\epsilon_1,n_1}B_{\epsilon_2,n_2}
     B_{\epsilon_3,n_3}B_{\epsilon_4,n_4}:
  }{\sqrt{n_1n_2n_3n_4}},
  \qquad
  B_{+1,n}=a_n,\quad B_{-1,n}=a_n^\dagger .
  \label{eq:phi4-dlcq-quartic-operator}
\end{equation}
The Kronecker condition is longitudinal momentum conservation.  If one uses
the field normalization
\[
  \phi(x^-)
  =
  \sum_{n>0}
  \frac{a_ne^{-\ii n x^-/R}+a_n^\dagger e^{\ii n x^-/R}}{\sqrt{4\pi n}},
  \qquad x^-\sim x^-+2\pi R,
\]
then \(g=\lambda/(96\pi)\) for the interaction
\(\lambda\int:\phi^4:/4!\) when written in invariant-mass units.  The
chapter treats \(g\) as the finite regulator coordinate, because a
cutoff-dependent coupling and possible zero-mode counterterms are part of a
continuum scheme.

The finite DLCQ invariant-mass matrix is
\begin{equation}
  M^2_K
  =
  M^2_{0,K}
  +
  gK\,P_KQ_KP_K ,
  \label{eq:phi4-dlcq-finite-mass-matrix}
\end{equation}
where \(P_K\) is the projection onto the finite partition basis
\eqref{eq:phi4-dlcq-partition-basis}.  Optional particle-number cutoffs may
be added as further finite-regulator coordinates, but they must then be
removed or extrapolated separately from the harmonic resolution.
\end{construction}

The matrix \eqref{eq:phi4-dlcq-finite-mass-matrix} is Hermitian.  The
coefficient in \eqref{eq:phi4-dlcq-quartic-operator} is real, and taking the
adjoint sends every ordered monomial with signs \(\epsilon_i\) to the
corresponding monomial with signs \(-\epsilon_i\).  The constraint
\(\sum_i\epsilon_i n_i=0\) is invariant under this sign change, and normal
ordering restores the displayed ordered form.  Since \(P_K\) is an
orthogonal projection onto a fixed total-resolution sector, the projected
finite matrix is Hermitian.

A small matrix element fixes the convention.  At \(K=3\), the states
\[
  \ket{1,1,1}=\frac{(a_1^\dagger)^3}{\sqrt{3!}}\ket0,
  \qquad
  \ket{3}=a_3^\dagger\ket0
\]
both lie in the same harmonic-resolution sector.  The only term in \(Q_K\)
that connects them is the ordered family with one \(a_3^\dagger\) and three
\(a_1\)'s, together with its four possible positions in the quartic product.
Thus
\begin{equation}
  \bra3 Q_3\ket{1,1,1}
  =
  \frac{4}{\sqrt3}\,\sqrt{3!}
  =
  4\sqrt2 .
  \label{eq:phi4-dlcq-one-three-check}
\end{equation}
For the three-particle state itself, the two-creation two-annihilation
normal-ordered term with all momenta equal to one gives
\begin{equation}
  \bra{1,1,1}Q_3\ket{1,1,1}
  =
  6\,\bra{1,1,1}(a_1^\dagger)^2a_1^2\ket{1,1,1}
  =
  36.
  \label{eq:phi4-dlcq-three-particle-diagonal-check}
\end{equation}
These finite identities are not physical mass predictions; they are
normalization checks for the regulator matrix.

The script
\[
  \texttt{qft\_scripts/phi4\_dlcq.py}
\]
implements Construction~\ref{con:finite-scalar-phi4-dlcq-matrix}.  Its smoke
mode builds a small positive-\(p^+\) matrix, checks finite Hermiticity, and
prints the lowest finite invariant masses.  The companion calculation check
\[
  \texttt{calculation-checks/hamiltonian\_truncation\_dlcq\_checks.py}
\]
verifies total harmonic resolution, zero-coupling reduction to
\eqref{eq:phi4-dlcq-free-mass}, Hermiticity, and the two convention-sensitive
matrix elements \eqref{eq:phi4-dlcq-one-three-check} and
\eqref{eq:phi4-dlcq-three-particle-diagonal-check}.

\section{\texorpdfstring{Large-\(N\) Two-Dimensional QCD: A Finite DLCQ Matrix}{Large-N Two-Dimensional QCD: A Finite DLCQ Matrix}}

The standard large-\(N\) two-dimensional QCD meson equation involves the
principal-value operator
\[
  (\mathcal K\phi)(x)
  =
  \operatorname{PV}\int_0^1
  \frac{\phi(x)-\phi(y)}{(x-y)^2}\,\dd y ,
  \qquad 0<x<1,
\]
together with endpoint mass terms
\[
  \frac{m_1^2}{x}\phi(x)+\frac{m_2^2}{1-x}\phi(x).
\]
This section records a finite regulator, not a continuum theorem.  Fix a
harmonic resolution \(K\geq3\) and grid points \(x_n=n/K\),
\(n=1,\ldots,K-1\).  For \(\gamma>0\), define the real symmetric matrix
\[
  (M_K^2)_{nm}
  =
  \delta_{nm}
  \left[
    \frac{m_1^2}{x_n}
    +
    \frac{m_2^2}{1-x_n}
    +
    \gamma K\sum_{\substack{j=1\\ j\neq n}}^{K-1}
      \frac1{(n-j)^2}
  \right]
  -
  (1-\delta_{nm})\frac{\gamma K}{(n-m)^2}.
\]
The normalization of \(\gamma\) depends on the convention for
\(g^2N/\pi\).  The finite object above is sufficient to illustrate harmonic
resolution, positivity, and extrapolation obligations.

\paragraph{Finite principal-value quadratic form.}
For \(m_1^2,m_2^2,\gamma>0\) and any
\(v=(v_1,\ldots,v_{K-1})\in\mathbb R^{K-1}\),
\begin{equation}
  v^T M_K^2 v
  =
  \sum_{n=1}^{K-1}
  \left(
    \frac{m_1^2}{x_n}
    +
    \frac{m_2^2}{1-x_n}
  \right)v_n^2
  +
  \gamma K
  \sum_{1\leq n<m\leq K-1}
  \frac{(v_n-v_m)^2}{(n-m)^2}.
  \label{eq:thooft-dlcq-finite-quadratic-form}
\end{equation}
In particular \(M_K^2\) is positive definite.

Substitute the definition of \(M_K^2\) into \(v^TM_K^2v\).  The mass terms
give the first displayed sum.  The off-diagonal kernel gives
\[
  \gamma K
  \sum_{n=1}^{K-1}
  \sum_{\substack{m=1\\m\neq n}}^{K-1}
  \frac{v_n^2-v_nv_m}{(n-m)^2}.
\]
Pair the term \((n,m)\) with \((m,n)\).  For each unordered pair
\(\{n,m\}\),
\[
  \frac{v_n^2-v_nv_m}{(n-m)^2}
  +
  \frac{v_m^2-v_mv_n}{(m-n)^2}
  =
  \frac{(v_n-v_m)^2}{(n-m)^2}.
\]
This gives the stated quadratic form.  Since \(0<x_n<1\), the mass
coefficients are positive; hence \(v^TM_K^2v>0\) for \(v\neq0\).
The calculation is a finite-regulator algebra check: it proves positivity of
the displayed matrix.  Positivity and self-adjointness of the continuum
principal-value operator require a separate continuum-domain construction.

The script
\[
  \texttt{qft\_scripts/thooft\_dlcq.py}
\]
constructs this matrix and diagonalizes it at finite \(K\).  A continuum
claim would require an analysis of the \(K\to\infty\) limit, endpoint
behavior, zero modes, normalization of \(\gamma\), and comparison with a
Lorentz-covariant reconstruction.  The script therefore prints finite
eigenvalues and a quadratic-form positivity check rather than presenting the
numbers as physical meson masses.

\section{Finite-Regulator Extrapolation}
\label{sec:finite-regulator-extrapolation}

The output of a lattice, Hamiltonian-truncation, or DLCQ script is a number
attached to a regulator.  A continuum claim is a separate mathematical
statement about a limiting procedure.  This section records the minimal
finite-dimensional facts used in the companion scripts and explains what an
extrapolation hypothesis must contain.

\begin{definition}[Finite-regulator observable record]
\label{def:finite-regulator-observable-datum}
A finite-regulator observable record consists of:
\[
  \bigl(I,\preceq;\ X_i,\ A_i,\ \sigma_i,\ \mathcal E_i\bigr) .
\]
Here \((I,\preceq)\) is a directed set of regulator labels, \(X_i\) is the
finite object being computed at label \(i\), \(A_i\in\mathbb R\) is the
finite observable extracted from \(X_i\), \(\sigma_i\ge0\) is the statistical
standard error if \(A_i\) is estimated by a random algorithm, and
\(\mathcal E_i\) is the declared systematic error model.  For deterministic
TCSA or DLCQ diagonalization one has \(\sigma_i=0\), but \(\mathcal E_i\) is
not zero: it contains cutoff, finite-volume, counterterm, endpoint, or
zero-mode errors.  For a Markov-chain computation, \(A_i\) itself is an
estimator and \(\sigma_i\) must include autocorrelation.
\end{definition}

\begin{remark}[Interpolation ambiguity of finite cutoff data]
\label{rem:finite-data-do-not-determine-continuum-limit}
Let \(K_1,\ldots,K_N\) be distinct positive regulator values and let
\(A_1,\ldots,A_N\in\mathbb R\) be finite outputs.  For every proposed
continuum value \(A_\infty\in\mathbb R\), there is a polynomial \(p(t)\) of
degree at most \(N\) such that
\[
  p(0)=A_\infty,
  \qquad
  p(K_j^{-1})=A_j,\quad j=1,\ldots,N .
\]
Consequently finitely many cutoff values, by themselves, cannot prove a
continuum limit or its value.
The \(N+1\) points
\[
  0,\quad K_1^{-1},\ldots,K_N^{-1}
\]
are distinct.  Lagrange interpolation gives a unique polynomial of degree at
most \(N\) taking the prescribed values \(A_\infty,A_1,\ldots,A_N\) at those
points.  Since \(A_\infty\) was arbitrary, the same finite data are
compatible with arbitrarily many candidate limits unless additional
information is supplied.  The obstruction is the missing analytic ansatz,
monotonicity estimate, remainder bound, or convergence theorem that would
restrict the allowed interpolating functions near \(K^{-1}=0\).
\end{remark}

\begin{definition}[Power-law cutoff expansion with a remainder bound]
\label{def:power-law-cutoff-expansion}
Let \(K\) be a positive regulator parameter tending to infinity.  A
deterministic sequence \(A_K\) has an \(m\)-term power-law cutoff expansion
with base exponent \(\omega>0\) and controlled remainder if there are real
numbers \(A_\infty,c_1,\ldots,c_m\), a constant \(C\), and \(K_0\) such that
\begin{equation}
  \left|
    A_K
    -
    A_\infty
    -
    \sum_{a=1}^m c_a K^{-a\omega}
  \right|
  \le
  C K^{-(m+1)\omega},
  \qquad K\ge K_0 .
  \label{eq:power-law-cutoff-remainder-bound}
\end{equation}
The same definition applies to a subsequence, for example
\(K,Ks_1,\ldots,Ks_m\), provided these are admissible regulator labels and
the displayed bound holds on that subsequence.  In a TCSA computation the
exponent \(\omega\) and the counterterm coordinates should be derived from
the high-energy tail or OPE; in a DLCQ computation they should be derived
from endpoint and zero-mode analysis.  A fit without
\eqref{eq:power-law-cutoff-remainder-bound} is a diagnostic, not a theorem.
\end{definition}

For a two-cutoff Richardson extrapolant, the leading correction cancels only
under a stated asymptotic expansion with a controlled remainder.  Assume
\[
  A_K=A_\infty+cK^{-\omega}+r_K,
  \qquad
  |r_K|\le C K^{-\eta},
  \qquad
  \eta>\omega>0 ,
\]
for \(K\ge K_0\).  Define
\begin{equation}
  R_K
  =
  \frac{2^\omega A_{2K}-A_K}{2^\omega-1}.
  \label{eq:two-cutoff-richardson-estimator}
\end{equation}
Then
\begin{equation}
  |R_K-A_\infty|
  \le
  \frac{2^\omega 2^{-\eta}+1}{2^\omega-1}\,
  C K^{-\eta}.
  \label{eq:two-cutoff-richardson-bound}
\end{equation}
Substitute the assumed expansion into
\eqref{eq:two-cutoff-richardson-estimator}:
\[
  2^\omega A_{2K}-A_K
  =
  (2^\omega-1)A_\infty
  +
  2^\omega c(2K)^{-\omega}
  -
  cK^{-\omega}
  +
  2^\omega r_{2K}-r_K .
\]
The two terms proportional to \(c\) cancel because
\(2^\omega(2K)^{-\omega}=K^{-\omega}\).  Dividing by \(2^\omega-1\) gives
\[
  R_K-A_\infty
  =
  \frac{2^\omega r_{2K}-r_K}{2^\omega-1}.
\]
Using \(|r_{2K}|\le C(2K)^{-\eta}\) gives
\eqref{eq:two-cutoff-richardson-bound}.

\paragraph{Integer-power extrapolation with explicit weights.}
Let \(s_0,\ldots,s_m\) be distinct positive real numbers and suppose
\(A_K\) satisfies \eqref{eq:power-law-cutoff-remainder-bound} on the
subsequence \(s_jK\), \(j=0,\ldots,m\).  Let \(w_0,\ldots,w_m\) be the unique
solution of
\begin{equation}
  \sum_{j=0}^m w_j=1,
  \qquad
  \sum_{j=0}^m w_j s_j^{-a\omega}=0,
  \quad a=1,\ldots,m .
  \label{eq:extrapolation-weight-moments}
\end{equation}
Then the extrapolated value
\[
  R_K^{(m)}
  =
  \sum_{j=0}^m w_j A_{s_jK}
\]
satisfies
\begin{equation}
  |R_K^{(m)}-A_\infty|
  \le
  C K^{-(m+1)\omega}
  \sum_{j=0}^m |w_j|\,s_j^{-(m+1)\omega}.
  \label{eq:integer-power-extrapolation-bound}
\end{equation}

Set \(q_j=s_j^{-\omega}\).  The numbers \(q_0,\ldots,q_m\) are distinct,
so the Vandermonde matrix \((q_j^a)_{0\le a,j\le m}\) is invertible.  Hence
\eqref{eq:extrapolation-weight-moments} has a unique solution.  Insert
\eqref{eq:power-law-cutoff-remainder-bound} into \(R_K^{(m)}\).  The
coefficient of \(A_\infty\) is \(\sum_jw_j=1\).  For
\(a=1,\ldots,m\), the coefficient of \(c_aK^{-a\omega}\) is
\(\sum_jw_js_j^{-a\omega}=0\).  The remaining error is bounded by the
triangle inequality:
\[
  \left|\sum_{j=0}^m w_j r_{s_jK}\right|
  \le
  \sum_{j=0}^m |w_j|\,C(s_jK)^{-(m+1)\omega}.
\]
This is \eqref{eq:integer-power-extrapolation-bound}.
The finite algebra is exact, but its use in a continuum claim is only as good
as the expansion and remainder estimate in
Definition~\ref{def:power-law-cutoff-expansion}.  The weights cancel declared
integer powers; the QFT problem separately supplies the evidence that these
powers are the correct cutoff expansion.

\paragraph{\texorpdfstring{Large-\(K\) diagnostic for the finite 't Hooft matrices}{Large-K diagnostic for the finite 't Hooft matrices}.}
For the finite matrices \(M_K^2\) whose quadratic form is displayed in
\eqref{eq:thooft-dlcq-finite-quadratic-form}, write their eigenvalues in
nondecreasing order as
\[
  \lambda_0(K)\leq\lambda_1(K)\leq\cdots\leq\lambda_{K-2}(K).
\]
For a fixed label \(\ell\), the numerical sequence
\[
  A_K=\lambda_\ell(K)
\]
is a finite-regulator sequence in the sense of
Definition~\ref{def:finite-regulator-observable-datum}.  Keeping the label
\(\ell\) fixed is only a diagnostic convention until one proves that the
corresponding continuum state is isolated and that nearby finite-\(K\) levels
do not interchange in a way that changes the limiting object.  Choose distinct
resolutions \(K_1<\cdots<K_N\), a base exponent \(\omega>0\), and an integer
\(p<N\).  Set
\[
  q_r=K_r^{-\omega},\qquad
  X_{ra}=q_r^a,\qquad
  r=1,\ldots,N,\quad a=0,\ldots,p .
\]
The least-squares polynomial fit is
\begin{equation}
  \widehat c=(X^T X)^{-1}X^T A,\qquad
  A=(A_{K_1},\ldots,A_{K_N})^T ,
  \label{eq:dlcq-large-K-least-squares-fit}
\end{equation}
provided \(X\) has full column rank.  The reported continuum diagnostic is
\(\widehat c_0\), the fitted intercept at \(q=0\).

\paragraph{\texorpdfstring{Finite large-\(K\) fit with an explicit remainder amplifier}{Finite large-K fit with an explicit remainder amplifier}.}
Assume that the finite data used in
\eqref{eq:dlcq-large-K-least-squares-fit} have the form
\begin{equation}
  A_{K_r}
  =
  A_\infty+\sum_{a=1}^p c_a K_r^{-a\omega}+\rho_r ,
  \qquad r=1,\ldots,N ,
  \label{eq:large-K-fit-model-with-remainder}
\end{equation}
with \(X\) of full column rank.  Let
\[
  B=(X^T X)^{-1}X^T .
\]
Then the fitted intercept satisfies
\begin{equation}
  |\widehat c_0-A_\infty|
  \leq
  \left(\sum_{r=1}^N |B_{0r}|\right)\max_{1\leq r\leq N}|\rho_r| .
  \label{eq:large-K-intercept-remainder-amplification}
\end{equation}
In particular, if all \(\rho_r\) vanish, the fit recovers \(A_\infty\)
exactly, independently of the overdetermination \(N>p+1\).

Let
\[
  c=(A_\infty,c_1,\ldots,c_p)^T,
  \qquad
  \rho=(\rho_1,\ldots,\rho_N)^T .
\]
Equation~\eqref{eq:large-K-fit-model-with-remainder} is the vector identity
\[
  A=Xc+\rho .
\]
Since \(X\) has full column rank, \(BX=I_{p+1}\).  Hence
\[
  \widehat c
  =
  BA
  =
  B(Xc+\rho)
  =
  c+B\rho .
\]
Taking the zeroth component gives
\[
  \widehat c_0-A_\infty
  =
  \sum_{r=1}^N B_{0r}\rho_r .
\]
The triangle inequality gives
\eqref{eq:large-K-intercept-remainder-amplification}.  If \(\rho=0\), then
\(\widehat c=c\) and the intercept is exact.

The companion script
\[
  \texttt{qft\_scripts/thooft\_dlcq\_extrapolation.py}
\]
implements this finite procedure for the matrices \(M_K^2\): it prints the
finite eigenvalues, the fitted intercepts, the residuals, the condition number
of \(X\), and the factor
\(\sum_r |B_{0r}|\) appearing in
\eqref{eq:large-K-intercept-remainder-amplification}.  These quantities
measure how a chosen finite data set responds to a chosen ansatz.  They do not
prove that the large-\(K\) limit of the large-\(N\) QCD meson equation exists
or that the selected exponent \(\omega\) is correct.  Those conclusions require
separate endpoint and zero-mode estimates producing a bound of the form
\eqref{eq:large-K-fit-model-with-remainder}.

The companion calculation
\[
  \texttt{calculation-checks/numerical\_extrapolation\_checks.py}
\]
verifies the algebra in
\eqref{eq:two-cutoff-richardson-estimator}--\eqref{eq:integer-power-extrapolation-bound}
with exact rational arithmetic.  It does not verify that a particular QFT
observable obeys \eqref{eq:power-law-cutoff-remainder-bound}; that remains a
problem-specific analytic estimate.

The companion calculation
\[
  \texttt{calculation-checks/hamiltonian\_truncation\_dlcq\_checks.py}
\]
also verifies the exact-intercept and constant-sequence cases of
the finite large-\(K\) fit calculation above.

\paragraph{Correlated fits and window stability.}
Production lattice, tensor-network, and Hamiltonian-truncation data are rarely
independent scalar data.  A continuum extrapolation specification at a chosen fit
window \(W=\{i_1,\ldots,i_N\}\) therefore contains
\[
  A=(A_{i_1},\ldots,A_{i_N})^T,\qquad
  \Sigma_W,\qquad
  f_0,\ldots,f_p,\qquad
  R_1,\ldots,R_N .
\]
Here \(A_i\) are the finite-regulator outputs, \(\Sigma_W\) is the covariance
matrix of their statistical errors on the window, \(f_a(i)\) are the declared
fit functions with \(f_0(i)=1\), and \(R_r\) is a deterministic bound or
declared systematic envelope for the part of \(A_{i_r}\) not represented by
the chosen fit functions.  Put
\[
  X_{ra}=f_a(i_r),\qquad
  X\in\operatorname{Mat}_{N\times(p+1)}(\mathbb R).
\]
Assume in this paragraph that \(\Sigma_W\) is positive definite and that
\(X\) has full column rank.  If an estimated covariance matrix is singular,
the shrinkage, projection, or diagonal approximation used to replace it is a
new part of the fit specification, not an invisible numerical detail.

The correlated least-squares coordinate is
\begin{equation}
  \widehat c
  =
  (X^T\Sigma_W^{-1}X)^{-1}X^T\Sigma_W^{-1}A .
  \label{eq:correlated-fit-coordinate}
\end{equation}
Writing
\[
  G=X^T\Sigma_W^{-1}X,\qquad
  B=G^{-1}X^T\Sigma_W^{-1},
\]
one has \(BX=I_{p+1}\).  If the data on the window decompose as
\begin{equation}
  A=Xc+\rho+\epsilon,
  \qquad
  \mathbb E\epsilon=0,\qquad
  \mathbb E(\epsilon\epsilon^T)=\Sigma_W ,
  \label{eq:correlated-fit-data-decomposition}
\end{equation}
then
\begin{equation}
  \widehat c-c=B\rho+B\epsilon .
  \label{eq:correlated-fit-error-decomposition}
\end{equation}
Thus the intercept \(\widehat A_\infty=\widehat c_0\) has the exact finite
error decomposition
\[
  \widehat A_\infty-A_\infty
  =
  \sum_{r=1}^N B_{0r}\rho_r
  +
  \sum_{r=1}^N B_{0r}\epsilon_r .
\]
If \(|\rho_r|\leq R_r\), the deterministic fit-window systematic coordinate is
\begin{equation}
  S_W=\sum_{r=1}^N |B_{0r}|R_r .
  \label{eq:correlated-fit-systematic-coordinate}
\end{equation}
The stochastic variance of the intercept is
\begin{equation}
  \sigma_{\infty,W}^2
  =
  e_0^T B\Sigma_WB^Te_0
  =
  e_0^TG^{-1}e_0 ,
  \label{eq:correlated-fit-intercept-variance}
\end{equation}
where \(e_0=(1,0,\ldots,0)^T\).  The second equality follows from
\[
  B\Sigma_WB^T
  =
  G^{-1}X^T\Sigma_W^{-1}\Sigma_W\Sigma_W^{-1}XG^{-1}
  =
  G^{-1}.
\]
The correlated residual
\begin{equation}
  \chi_W^2
  =
  (A-X\widehat c)^T\Sigma_W^{-1}(A-X\widehat c)
  \label{eq:correlated-fit-chi-square-coordinate}
\end{equation}
is a finite diagnostic for the chosen window and model.  Its probabilistic
interpretation requires the sampling law of the estimator, the covariance
estimator, and the treatment of autocorrelation described in
Section~\ref{sec:blocking-jackknife-bootstrap-error-coordinates}.

A window-stability analysis is the finite collection
\[
  \mathcal W=\{(W_\alpha,X_\alpha,\Sigma_\alpha,R_\alpha)\}_{\alpha\in A}
\]
together with the resulting coordinates
\[
  \widehat A_{\infty,\alpha},\qquad
  \sigma_{\infty,\alpha},\qquad
  S_\alpha,\qquad
  \chi_\alpha^2 .
\]
Agreement of these coordinates across windows is evidence about the chosen
finite-regulator family only after the windows use the same physical
observable, the same renormalization convention, and compatible operator
normalizations.  The analytic estimate producing the envelopes
\(R_\alpha\) remains the source of continuum control.  For tensor-network and
neural-state calculations the window label must also record bond dimension or
architecture, optimizer termination criterion, local-energy variance or
discarded weight, and any symmetry projection used to select the finite
sector.  Otherwise a stable sequence of numbers may be comparing different
finite problems rather than a single regulated QFT observable.

The reader-facing implementation
\[
  \texttt{qft\_scripts/finite\_regulator\_extrapolation.py}
\]
takes this paragraph literally.  Its input data consist of regulator labels,
finite observable values, an optional covariance matrix, nonnegative
remainder envelopes, and half-open fit windows.  For each window it constructs
\[
  X_{ra}=K_r^{-a\omega},\qquad
  B=(X^T\Sigma_W^{-1}X)^{-1}X^T\Sigma_W^{-1},
\]
then reports \(\widehat A_{\infty,W}\), the propagated statistical error
\((e_0^TG^{-1}e_0)^{1/2}\), the deterministic coordinate
\(\sum_r |B_{0r}|R_r\), the correlated residual, the condition number, and
the spread of fitted intercepts across the declared windows.  Supplying the
identity matrix as a covariance, or using diagonal errors, is part of the
finite fit specification and is printed as such.  The script therefore checks the
bookkeeping of a proposed extrapolation.  The cutoff exponent, counterterm
scheme, and remainder estimate are mathematical inputs whose justification
comes from the regulated theory, not from the fit output alone.

\section{Light-Front Hamiltonian}

The light-front Hamiltonian convention used in this chapter is the regulator
specification of Definition~\ref{def:light-front-hamiltonian-regulator-datum}.  The
finite numerical sections above use its DLCQ specialization; the analytic
large-\(N\) QCD construction in
Chapter~\ref{ch:large-n-two-dimensional-qcd-light-front} and the
Chern--Simons--matter light-cone construction in
Chapter~\ref{ch:susy-three-dimensional-chern-simons-matter} use the same
coordinate convention, but with model-specific constraint equations instead
of a universal finite-matrix recipe.

\section{Cross-Method Benchmark Comparisons}
\label{sec:cross-method-numerical-benchmark-comparisons}

Lattice Monte Carlo, Hamiltonian truncation, TCSA/TFFSA, and DLCQ produce
different finite mathematical objects.  A numerical agreement between their
outputs is meaningful only after a common QFT observable and a common
normalization have been specified.  The comparison itself is therefore part
of the regulated calculation.

\begin{definition}[Cross-method benchmark specification]
\label{def:cross-method-benchmark-datum}
A cross-method benchmark specification consists of the following information.
\begin{enumerate}
  \item A target QFT observable \(Q\).  Examples are a dimensionless mass
  ratio, a finite-volume energy gap in units of the physical mass, a matrix
  element normalized by a declared two-point function, or a Euclidean
  distribution paired with a specified test function.
  \item A finite set of methods \(m\in\mathcal M\).  For each method one
  specifies the finite regulator object \(X_m\): for Euclidean lattice Monte
  Carlo this includes lattice spacing, finite volume, boundary conditions,
  action, update kernel, and measured estimator; for Hamiltonian truncation
  it includes the finite Hilbert space, Hamiltonian matrix, counterterm
  coordinates, symmetry sector, and extraction map; for DLCQ it includes
  harmonic resolution, zero-mode prescription, constraint sector, and finite
  invariant-mass matrix.
  \item A normalization map
  \[
    N_m:X_m\longrightarrow\mathbb R^r
  \]
  which converts the method-specific finite output to the same coordinate
  system for \(Q\).  Field-strength conventions, finite-volume units,
  operator normalizations, and sector labels are part of \(N_m\).
  \item Statistical error coordinates \(s_m\), deterministic finite-regulator
  envelopes \(R_m\), and a scheme-matching coordinate \(b_m\) measuring any
  residual mismatch between the method's finite observable and the declared
  target observable.
  \item A limiting or extrapolation hypothesis, when a continuum statement is
  claimed, stating that
  \[
    N_m(X_m)=Q+\epsilon_m,\qquad
    |\epsilon_m|\leq s_m+R_m+b_m
  \]
  in the declared norm or componentwise in \(\mathbb R^r\).
\end{enumerate}
\end{definition}

The comparison check between two methods is then finite arithmetic.  If
two methods \(m,n\) are claimed to compute the same coordinate of \(Q\), their
finite outputs must satisfy
\begin{equation}
  |N_m(X_m)-N_n(X_n)|
  \leq
  s_m+s_n+R_m+R_n+b_m+b_n
  \label{eq:cross-method-finite-consistency-bound}
\end{equation}
for the claimed error coordinates.  Failure of
\eqref{eq:cross-method-finite-consistency-bound} means that at least one of
the finite estimates, error models, normalization maps, or target
identifications is wrong.  Satisfaction of
\eqref{eq:cross-method-finite-consistency-bound} supplies a finite
compatibility check conditional on the stated extrapolation hypotheses.
The continuum statement is the separate proof that the hypotheses and
envelopes follow from the regulated QFT.

The main finite objects in the present numerical layer are as follows.
\begin{description}[leftmargin=0.24\textwidth,style=nextline]
  \item[Euclidean lattice Monte Carlo.]
  The finite object is a positive measure, a Markov kernel, and sampled
  correlators.  The usual spectral coordinate is a transfer-matrix gap
  extracted from a correlator matrix or effective mass.  The continuum burden
  is reflection positivity, volume control, autocorrelation analysis,
  continuum scaling, and operator matching.
  \item[Hamiltonian truncation/TCSA/TFFSA.]
  The finite object is a sector of a Hilbert space together with a projected
  Hamiltonian matrix.  The usual spectral coordinates are Ritz eigenvalues,
  residual bounds, and finite matrix elements.  The continuum burden is
  counterterm determination, cutoff convergence, finite-volume control, and
  operator normalization.
  \item[DLCQ/light-front truncation.]
  The finite object is a fixed harmonic-resolution physical sector together
  with a finite \(M^2\) matrix.  The usual spectral coordinate is an ordered
  invariant-mass eigenvalue at fixed \(K\).  The continuum burden is
  zero-mode constraints, endpoint behavior, large-\(K\) estimates, and
  Lorentz-covariant reconstruction.
\end{description}

\begin{definition}[Benchmark manifest]
\label{def:benchmark-manifest}
A benchmark manifest is a finite, machine-readable representative of
Definition~\ref{def:cross-method-benchmark-datum}.  It consists of:
\begin{enumerate}
  \item the target observable label \(Q\), the dimension \(r\) of its reported
  coordinate vector, and the norm in which the comparison is to be made;
  \item for each method \(m\), a regulator description, a normalization
  identifier, the coordinate vector \(x_m\in\mathbb R^r\), the componentwise
  statistical error \(\sigma_m\), regulator envelope \(\rho_m\), and
  matching error \(\beta_m\);
  \item optional covariance data for the statistical part, with the
  convention that covariance data never replace the declared deterministic
  envelopes \(\rho_m,\beta_m\);
  \item provenance data: script name or analysis notebook, input dataset,
  random seed where relevant, regulator parameters, and any hash or version
  identifier needed to reproduce the finite number \(x_m\).
\end{enumerate}
The declared componentwise error vector is
\begin{equation}
  e_m=\sigma_m+\rho_m+\beta_m.
  \label{eq:benchmark-manifest-total-error}
\end{equation}
The manifest is consistent at the finite-regulator level when, for every pair
of methods \(m,n\),
\begin{equation}
  |x_m-x_n|_{\rm comp}\leq e_m+e_n,
  \label{eq:benchmark-manifest-pairwise-test}
\end{equation}
where the absolute value and inequality are componentwise.  The script
\path{qft_scripts/benchmark_manifest_consistency.py} implements exactly
\eqref{eq:benchmark-manifest-pairwise-test} for JSON manifests.  A passing
manifest verifies only that the finite reported coordinates are mutually
compatible with the declared error envelopes and normalization identifiers.
It does not verify that the envelopes are correct, that the regulator limits
exist, that the methods compute the same continuum QFT observable, or that
the target observable is nonperturbatively constructed.
\end{definition}

This manifest discipline is intentionally stricter than a table of numerical
values.  If a lattice correlator fit and a Hamiltonian-truncation
diagonalization report the same mass ratio but use different finite-volume
normalizations, the mismatch belongs in \(\beta_m\) or in a failed
normalization identifier before any comparison is made.  If two finite
coordinates disagree beyond \eqref{eq:benchmark-manifest-pairwise-test}, the
failure has a definite meaning: at least one coordinate, uncertainty model,
normalization map, or target identification is not the one recorded in the
manifest.  If the manifest passes, the continuum conclusion still begins
only after the separate estimates on \(\sigma_m,\rho_m,\beta_m\) are proved
or justified in the relevant regulator framework.

The companion scripts instantiate this comparison at finite regulator.  The
Euclidean side contains finite Ising, scalar \(\phi^4\), \(\mathbb Z_2\),
\(SU(2)\), and \(SU(3)\) samplers, plus autocorrelation,
static-potential, Wilson-flow, smearing, topological-charge, HMC/RHMC, and
glueball-GEVP diagnostics.  The Hamiltonian side contains exactly solvable
Ising thermal truncation, compact-boson sine-Gordon vertex assembly,
nonintegrable scalar \(\phi^4\) truncation, connected Ising TFFSA blocks,
spectral-flow diagnostics, and the exact \(E_8\) target table.  The
light-front side contains scalar \(\phi^4\) DLCQ, the large-\(N\)
two-dimensional QCD matrix, and a large-\(K\) extrapolation diagnostic.  The
generic finite-regulator extrapolation script records the common
window/covariance/remainder-envelope data used by all three classes of
methods, while the benchmark-manifest script records pairwise compatibility
of normalized outputs across methods.

\phantomsection
\subsection*{Public Finite-Regulator Check Index}
\label{subsec:public-finite-regulator-check-index}

The companion-script layer is part of the mathematical presentation only
when each script is read through a finite verification check.  A smoke run by itself
is an implementation regression; it is not a QFT result.  The finite
check attached to a script is the tuple
\[
  \mathcal C_{\rm num}
  =
  \bigl(
    \mathcal D_{\rm fin},
    F_{\rm fin},
    \mathcal I_{\rm check},
    \mathcal S_{\rm scope}
  \bigr),
\]
where \(\mathcal D_{\rm fin}\) is the declared finite regulator specification,
\(F_{\rm fin}\) is the finite output map computed by the script,
\(\mathcal I_{\rm check}\) is the list of algebraic, spectral, measure, or
statistical identities checked against the chapter, and
\(\mathcal S_{\rm scope}\) is the boundary between the finite assertion and
any continuum or infinite-volume assertion.  The smoke harness checks that
the implementation still produces its declared finite output.  The
corresponding calculation-check scripts, when present, check the identities
in \(\mathcal I_{\rm check}\).  Neither layer replaces the estimates needed
to pass from \(\mathcal D_{\rm fin}\) to a continuum QFT observable.

The public scripts currently have the following finite-check meanings.
\begin{description}[leftmargin=2em,labelindent=0pt,labelsep=0.5em,style=nextline]
\item[Spin and scalar Metropolis.]
The Ising and scalar \(\phi^4\) Metropolis scripts compute finite periodic
lattice expectations for explicitly declared Boltzmann weights.  Their
check is invariance of the finite target measure under the displayed
single-site Metropolis kernels, together with local energy or
action-difference identities.  They do not prove Onsager scaling,
constructive \(\phi^4_2\), a critical surface, or an Ising CFT limit.

\item[Compact gauge samplers.]
The finite \(\mathbb Z_2\), \(SU(2)\) Metropolis, \(SU(2)\)
heat-bath and overrelaxation, and \(SU(3)\) HDF5 sampler scripts compute finite
compact link-integrals and Wilson-loop estimators.  Their checks are
gauge invariance of the measured finite observables, Haar measure
preservation, detailed balance or conditional heat-bath invariance, and the
declared data coordinate convention.  They do not prove continuum
Yang--Mills, mass gaps, confinement, topological sector sampling, or a
production continuum extrapolation.

\item[Flow, smearing, topology, and static-potential analysis.]
The Wilson flow, APE smearing, clover topology diagnostic,
static potential, and glueball GEVP scripts compute finite analysis
functionals of saved lattice data.  Their checks are finite group
projection, covariance of smearing, the declared clover diagnostic, nonlinear
jackknife and bootstrap propagation for Wilson-loop ratios, and finite
generalized eigenvalue residual control.  They do not define a continuum
topological charge, a continuum static potential, or a physical glueball
spectrum without the continuum estimates of
Chapters~\ref{ch:constructive-wilson-lattice-gauge-theory} and
\ref{ch:constructive-lattice-continuum-local-qft}.

\item[HMC, RHMC, and Markov-chain error analysis.]
The HMC and RHMC smoke module and the autocorrelation and resampling script compute
finite algorithmic diagnostics: HMC reversibility defects, Hamiltonian
changes, positive-matrix spectral edges, rational pseudofermion actions,
solver residuals, autocorrelations, block errors, and jackknife and bootstrap
coordinates.  Their checks are the finite symplectic and reversibility
identities, determinant or rational-determinant bookkeeping, and finite
time-series formulae.  They do not prove ergodic sampling of a large
production ensemble or a continuum-determinant limit.

\item[Hamiltonian truncation and TCSA/TFFSA matrices.]
The Ising energy benchmark, sine-Gordon zero-mode, sine-Gordon oscillator
vertex, scalar \(\phi^4\) truncation, connected Ising TFFSA,
TFFSA spectral-flow, and \(E_8\) target-data scripts compute finite matrices
or exact finite target data.  Their checks are finite Hermiticity,
selection rules, normal-ordering conventions, residual or spectral-flow
identities, and the Perron--Frobenius convention check for the \(E_8\) target
table.  They do not prove cutoff-extrapolated spectra.  For the sine-Gordon,
nonintegrable \(\phi^4\), and magnetic-Ising TCSA/TFFSA \(E_8\) comparisons,
the counterterm, finite-volume, and cutoff-error data specified above remain
separate hypotheses.

\item[DLCQ and light-front finite matrices.]
The scalar \(\phi^4\) DLCQ, large-\(N\) two-dimensional QCD DLCQ, and
large-\(K\) extrapolation scripts compute finite matrices at harmonic
resolution \(K\), together with diagnostics at fixed spectral label.  Their
checks are positive longitudinal partition bookkeeping, finite
principal value or normal ordered matrix identities, positivity of the finite
quadratic form in the declared model, and finite least squares amplification
for a declared large-\(K\) ansatz.  They do not prove the \(K\to\infty\) continuum
spectrum without zero-mode, endpoint, constraint, and remainder estimates.

\item[Extrapolation, manifests, and cluster wrappers.]
The finite-regulator extrapolation, benchmark-manifest, and cluster-grid
scripts compute finite bookkeeping checks: correlated-window intercepts
with declared covariance and remainder envelopes, componentwise
benchmark-manifest compatibility, and deterministic grid-to-task assignment
for cluster sweeps.  They do not create a cutoff ansatz, validate a
systematic error model, or promote a small cluster run to a continuum QFT
calculation.
\end{description}

This index is deliberately redundant with the comments printed by the
scripts and with \path{qft_scripts/README.md}, where exact filenames and
smoke commands are listed.  Its role in the monograph is to make the finite
mathematical object visible before the reader interprets any number produced
by a public script.

For the magnetic Ising theory, the present public scripts deliberately stop
at finite checks: the TFFSA connected-block matrix fixes the spin-field
normalization convention, the spectral-flow script checks finite
Hellmann--Feynman slopes, and the \(E_8\) script fixes the continuum target
mass-ratio table.  A full comparison between a finite TCSA/TFFSA spectrum,
Euclidean lattice correlators, and the exact \(E_8\) masses still requires a
declared finite-volume convention, cutoff counterterms, extrapolation
envelopes, and comparison estimates in the sense of
Definition~\ref{def:cross-method-benchmark-datum}.

\section{Benchmark Protocol}

A benchmark for any truncation method should specify:
\[
\begin{gathered}
  \text{regulated Hilbert space},\qquad
  \text{operator basis},\qquad
  \text{counterterm coordinates},\\
  \text{observables used for matching},\qquad
  \text{extrapolation model},\qquad
  \text{independent checks}.
\end{gathered}
\]
Independent checks include Ward identities, known finite-volume limits,
reflection positivity where applicable, spectral sum rules, and comparison
with Euclidean lattice or exact integrable data in overlapping regimes.

The public companion-script policy is recorded in
\texttt{planning/14\_code\_policy.md}.  The script smoke harness
\[
  \texttt{tools/run\_qft\_scripts\_smoke.sh}
\]
runs the finite lattice, Hamiltonian-truncation, TCSA/TFFSA, DLCQ,
finite-extrapolation, and benchmark-manifest smoke examples.  Passing that
harness verifies only the finite computations and internal consistency
checks just described.

\begin{openproblem}[Controlled QFT truncation]
\label{op:controlled-qft-hamiltonian-truncation}
Develop Hamiltonian truncation and DLCQ schemes with computable error bounds
for interacting QFT observables, including counterterm determination,
constraint preservation, zero-mode treatment, continuum extrapolation, and
public benchmark scripts.
\end{openproblem}
