\chapter{Constructive Scalar Models and OS Data}
\label{ch:constructive-scalar-models-os-data}

\ConventionsEuclidean

Constructive scalar models are the basic interacting examples in which the
objects used throughout this monograph are produced by estimates rather than
postulated: Schwinger functions, reflection positivity, Euclidean covariance,
growth bounds, and the Hilbert-space fields reconstructed from them.  The
word ``constructive'' will always mean a theorem about a limiting object in a
specified topology.  A finite cutoff integral, by itself, is only an input.

This chapter focuses on the scalar line beginning with \(P(\phi)_2\) and
\(\Phi^4_3\).  It does not treat these theories as formal path integrals
with unspecified counterterms.  It records the regulated variables, the Wick
and local counterterm coordinates appearing in the single regulated action,
the estimates that control the infinite-volume and ultraviolet limits, and
the OS data produced at the end.

\section{Gaussian Reference Field and Wick Powers}

Let \(m>0\), let
\[
  C=(-\Delta+m^2)^{-1}
\]
be the massive Euclidean covariance on \(\R^d\), and let \(\mu_C\) be the
centered Gaussian generalized field with covariance
\begin{equation}
  \mathbb E_{\mu_C}[\phi(f)\phi(g)]
  =
  \int_{\R^d\times\R^d} f(x)C(x-y)g(y)\,\dd^dx\,\dd^dy .
  \label{eq:constructive-gaussian-covariance}
\end{equation}
The field \(\phi\) is not a pointwise function.  The expression
\(\phi(x)^k\) is therefore not defined until a regularization and a limiting
procedure have been supplied.

\begin{definition}[Wick powers by Gaussian regularization]
\label{def:constructive-wick-powers-gaussian-regularization}
Let \(\rho_\varepsilon\) be a smooth approximate identity and set
\(\phi_\varepsilon=\rho_\varepsilon\ast\phi\).  Write
\[
  C_\varepsilon(0)
  =
  \mathbb E_{\mu_C}[\phi_\varepsilon(x)^2],
\]
which is translation independent.  The Wick powers used below are the
Hermite polynomials with variance \(C_\varepsilon(0)\):
\begin{align}
  :\!\phi_\varepsilon^2\!:
    &= \phi_\varepsilon^2-C_\varepsilon(0),\\
  :\!\phi_\varepsilon^3\!:
    &= \phi_\varepsilon^3-3C_\varepsilon(0)\phi_\varepsilon,\\
  :\!\phi_\varepsilon^4\!:
    &= \phi_\varepsilon^4
       -6C_\varepsilon(0)\phi_\varepsilon^2
       +3C_\varepsilon(0)^2 .
       \label{eq:constructive-wick-four}
\end{align}
After smearing against a test function, the Wick power
\(:\!\phi^k\!:(f)\) is the \(L^p(\mu_C)\)-limit of
\(\int f(x):\!\phi_\varepsilon(x)^k\!:\,\dd^dx\) when that limit exists.
\end{definition}

\paragraph{Origin of the Wick coefficients.}
For a centered Gaussian variable \(X\) with variance \(c\), the Wick
polynomial is
\[
  :\!X^n\!:
  =
  c^{n/2}H_n(X/\sqrt c),
\]
where \(H_n\) is the probabilists' Hermite polynomial.  In particular,
\[
  :\!X^2\!:=X^2-c,\qquad
  :\!X^3\!:=X^3-3cX,\qquad
  :\!X^4\!:=X^4-6cX^2+3c^2.
\]
The generating function
\[
  \exp(tX-\tfrac12ct^2)
  =
  \sum_{n\geq0} \frac{t^n}{n!}:\!X^n\!:
\]
is characterized by the property that its expectation against
\(\exp(sX)\) equals
\[
  \mathbb E\!\left[
    \exp(tX-\tfrac12ct^2)\exp(sX)
  \right]
  =
  \exp(cst+\tfrac12cs^2).
\]
Thus the coefficient of \(t^n\) is orthogonal to all polynomials of degree
less than \(n\) and has leading term \(X^n\), which is the defining
orthogonality property of the Hermite polynomial.  Expanding the generating
function through order four gives the displayed coefficients.

The calculation-check companion
\path{calculation-checks/constructive_scalar_spde_checks.py} verifies
these coefficients and the cutoff-asymptotic coefficients used below.

\section{\texorpdfstring{Finite-Volume \(P(\phi)_2\)}{Finite-Volume P(phi)2}}

Let \(\Lambda\subset\R^2\) be a square or torus, and let
\[
  P(q)=\sum_{j=0}^{2r} p_j q^j,
  \qquad p_{2r}>0,
\]
be bounded below.  Wick ordering is always relative to the Gaussian
covariance used in the finite regulator.  The regulated interaction is
\begin{equation}
  V_{\varepsilon,\Lambda}(\phi)
  =
  \int_\Lambda :\!P(\phi_\varepsilon(x))\!:\,\dd^2x .
  \label{eq:p-phi-two-regulated-interaction}
\end{equation}
The finite-volume Schwinger functional is
\begin{equation}
  S_{\varepsilon,\Lambda,n}(f_1,\ldots,f_n)
  =
  \frac{1}{Z_{\varepsilon,\Lambda}}
  \int
  \prod_{i=1}^n\phi_\varepsilon(f_i)\,
  \exp[-V_{\varepsilon,\Lambda}(\phi)]\,\dd\mu_C(\phi).
  \label{eq:p-phi-two-schwinger-regulated}
\end{equation}
This formula is meaningful because the cutoff field is a finite-variance
Gaussian process after smearing.  The theorem is the existence and
identification of the limits in which \(\varepsilon\downarrow0\) and
\(\Lambda\uparrow\R^2\).

\begin{definition}[Negative Sobolev control norm]
\label{def:p-phi-two-negative-sobolev-norm}
Fix a bounded open set \(\Lambda_1\) containing \(\Lambda\) and the supports
reached by \(\rho_\varepsilon\ast f\) for all sufficiently small
\(\varepsilon\).  For \(s>0\), define
\begin{equation}
  \|\phi\|_{H^{-s}(\Lambda_1)}^2
  :=
  \sum_{\nu}
  (1+\lambda_\nu)^{-s}\,|\phi(e_\nu)|^2,
  \label{eq:p-phi-two-negative-sobolev-norm}
\end{equation}
where \(\{e_\nu\}\) is any smooth orthonormal eigenbasis of a positive
self-adjoint realization of \(-\Delta\) on \(\Lambda_1\) with eigenvalues
\(\lambda_\nu\).  Different standard boundary conditions give equivalent
norms on smaller compact subsets, so the estimate below is independent of
that auxiliary choice after the constants are changed.
\end{definition}

\begin{quotedtheorem}[Nelson--Glimm--Jaffe--Spencer stability in \(P(\phi)_2\)]
\label{qthm:p-phi-two-stability}
Let \(\Lambda\subset\R^2\) be bounded, let \(P(q)\) be a real polynomial of
even degree with positive leading coefficient and bounded below, and let
\[
  V_{\varepsilon,\Lambda}
  =
  \int_\Lambda :\!P(\phi_\varepsilon(x))\!:\,\dd^2x
\]
be Wick ordered with respect to the same massive covariance as the Gaussian
measure.  There are \(s>0\), \(0<\alpha<2\), constants
\(A_\Lambda,B_\Lambda<\infty\), and \(\varepsilon_0>0\), independent of
\(\varepsilon\in(0,\varepsilon_0]\), such that for \(\mu_C\)-almost every
field
\begin{equation}
  V_{\varepsilon,\Lambda}(\phi)
  \geq
  -A_\Lambda-B_\Lambda\|\phi\|_{H^{-s}(\Lambda_1)}^{\alpha}.
  \label{eq:p-phi-two-stability-estimate}
\end{equation}
Consequently, for every finite \(p\geq1\),
\begin{equation}
  \sup_{0<\varepsilon\leq\varepsilon_0}
  \mathbb E_{\mu_C}\!\left[
    \exp(-pV_{\varepsilon,\Lambda})
  \right]<\infty .
  \label{eq:p-phi-two-uniform-exp-integrability}
\end{equation}
Together with the Gaussian-chaos convergence of the Wick monomials, this
implies that \(\exp[-V_{\varepsilon,\Lambda}]\) converges in \(L^p(\mu_C)\),
for every finite \(p\), to the finite-volume \(P(\phi)_2\) density.
\end{quotedtheorem}

The quoted theorem is a compact formulation of the stability and exponential
integrability estimates used in the Nelson, Glimm--Jaffe--Spencer, and
Guerra--Rosen--Simon constructions of \(P(\phi)_2\).  Its proof is not the
pointwise observation that \(P\) is bounded below; pointwise lower bounds for
\(:\!P(X)\!:\) contain covariance-dependent negative constants that diverge
as \(\varepsilon\downarrow0\).  The constructive proof decomposes the
covariance into scales, estimates the Wick-contraction combinatorics in each
cell, absorbs lower-degree Wick terms into the positive leading part by
large-field regulators, and uses the two-dimensional logarithmic covariance
to obtain the subquadratic Sobolev exponent \(\alpha<2\).  The theorem is
therefore imported here as a quoted constructive estimate, not hidden inside
the formal proof below.

\begin{theorem}[Finite-volume ultraviolet limit in \(P(\phi)_2\)]
\label{thm:p-phi-two-finite-volume-uv-limit}
For bounded \(\Lambda\), bounded-below polynomial \(P\), and Schwartz test
functions supported in the interior of \(\Lambda\), the limits
\[
  S_{\Lambda,n}(f_1,\ldots,f_n)
  =
  \lim_{\varepsilon\downarrow0}
  S_{\varepsilon,\Lambda,n}(f_1,\ldots,f_n)
\]
exist after Wick ordering.  The limiting finite-volume measure is absolutely
continuous with respect to the massive Gaussian free field restricted to
\(\Lambda\), with density represented by
\(\exp[-\int_\Lambda :\!P(\phi(x))\!:\dd^2x]\) as an \(L^p\)-limit for every
finite \(p\).
\end{theorem}

\begin{proof}
The \(L^p\)-convergence of the smeared Wick monomials follows from the
covariance estimate
\[
  \mathbb E[
    :\!\phi_\varepsilon^k\!:(f)
    :\!\phi_{\varepsilon'}^k\!:(f)]
  =
  k!\int f(x)C_{\varepsilon,\varepsilon'}(x-y)^k f(y)\,\dd^2x\,\dd^2y ,
\]
where the right-hand side converges because \(C(x)\) has only logarithmic
short-distance singularity in two dimensions and \(C(x)^k\) is locally
integrable for each fixed \(k\).  Hypercontractivity of Gaussian chaoses
upgrades \(L^2\)-convergence to \(L^p\)-convergence for finite \(p\).

The nontrivial point is uniform integrability of the exponential density.
This is exactly the content of
Quoted Theorem~\ref{qthm:p-phi-two-stability}.  To see how it enters, choose
the Sobolev index \(s>0\) in
\eqref{eq:p-phi-two-negative-sobolev-norm}.  The massive Gaussian free field
restricted to \(\Lambda_1\) belongs to \(H^{-s}(\Lambda_1)\) almost surely,
because
\[
  \mathbb E_{\mu_C}\|\phi\|_{H^{-s}(\Lambda_1)}^2
  =
  \sum_\nu(1+\lambda_\nu)^{-s}
  \langle e_\nu,Ce_\nu\rangle
  <\infty
\]
for \(s>0\) in two dimensions; the last inequality uses the elliptic bound
\(\langle e_\nu,Ce_\nu\rangle\leq C_0(1+\lambda_\nu)^{-1}\) and Weyl's law
for \(-\Delta\) on a bounded planar region.  Fernique's theorem gives some
\(\eta>0\) with
\[
  \mathbb E_{\mu_C}\exp\{
    \eta\|\phi\|_{H^{-s}(\Lambda_1)}^2
  \}<\infty .
\]
Since \(\alpha<2\), Young's inequality gives, for every \(p<\infty\),
\[
  pB_\Lambda \|\phi\|_{H^{-s}}^\alpha
  \leq
  \eta\|\phi\|_{H^{-s}}^2+C_{p,\eta,\alpha,B_\Lambda}.
\]
Combining this with \eqref{eq:p-phi-two-stability-estimate} proves
\eqref{eq:p-phi-two-uniform-exp-integrability}.  The
\(L^q\)-convergence of the Wick polynomial integrals gives convergence of
\(V_{\varepsilon,\Lambda}\) in probability.  Applying
\eqref{eq:p-phi-two-uniform-exp-integrability} with an exponent strictly
larger than \(p\) gives uniform integrability of
\(\exp[-pV_{\varepsilon,\Lambda}]\), hence \(L^p\)-convergence of the
densities.  Finally, the smeared insertions
\(\phi_\varepsilon(f_i)\) converge in every Gaussian \(L^q\) space, and
Holder's inequality combines this convergence with the uniform exponential
moments.  Therefore the normalized Schwinger functions have the stated
finite-volume ultraviolet limits.  The argument is local in \(\Lambda\), so
test functions supported away from the boundary are unaffected by the
boundary condition used to define the finite covariance.
\end{proof}

\section{Cluster Expansion and Infinite Volume}

The ultraviolet limit at fixed \(\Lambda\) does not yet construct the
infinite-volume QFT.  For massive \(P(\phi)_2\) the large-volume limit is
controlled by a cluster expansion.  The following formulation isolates the
estimate that turns local finite-volume data into global Schwinger functions.

\begin{definition}[Polymer activity and cluster norm]
\label{def:constructive-polymer-activity}
Partition \(\R^2\) into unit cells \(\Delta\).  A polymer \(X\) is a finite
connected union of cells.  Fix a finite-range symmetric incompatibility
relation \(X\nsim Y\); in the cell decompositions used below this means that
the \(R_0\)-cell neighbourhoods of \(X\) and \(Y\) intersect, for a fixed
integer \(R_0\).  Write \(X\sim Y\) for compatibility.  A polymer activity
assigns to \(X\) a functional
\[
  K(X,\phi)
\]
depending only on the field near \(X\).  For constants \(a,b>0\), define the
numerical majorant
\begin{equation}
  \kappa(X)
  :=
  \sup_{\phi}
  \left(
    |K(X,\phi)|e^{-b\|\phi\|_{X}^{2r}}
  \right)
  \label{eq:constructive-polymer-majorant}
\end{equation}
and the
weighted cluster norm
\begin{equation}
  \|K\|_{a,b}
  =
  \sup_{\Delta_0}
  \sum_{X\ni\Delta_0}
  e^{a|X|}
  \kappa(X),
  \label{eq:constructive-polymer-norm}
\end{equation}
where \(\|\phi\|_X\) is the chosen local large-field norm.  The exact
constructive proof fixes the Sobolev regularity and large-field regulator;
\eqref{eq:constructive-polymer-norm} records the structural requirement.
\end{definition}

\begin{lemma}[Penrose tree-graph bound]
\label{lem:constructive-penrose-tree-graph-bound}
Let \(X_1,\ldots,X_n\) be polymers and set
\[
  u_{ij}:=
  \begin{cases}
    1,& X_i\nsim X_j,\\
    0,& X_i\sim X_j .
  \end{cases}
\]
For \(n\geq2\), define the hard-core Ursell coefficient
\begin{equation}
  \varphi_{\mathrm c}(X_1,\ldots,X_n)
  :=
  \sum_{G\in\mathcal C_n}
  \prod_{\{i,j\}\in E(G)}(-u_{ij}),
  \label{eq:constructive-ursell-coefficient}
\end{equation}
where \(\mathcal C_n\) is the set of connected simple graphs on
\(\{1,\ldots,n\}\).  For \(n=1\), set \(\varphi_{\mathrm c}(X_1)=1\).
Then
\begin{equation}
  |\varphi_{\mathrm c}(X_1,\ldots,X_n)|
  \leq
  \sum_{T\in\mathcal T_n}
  \prod_{\{i,j\}\in E(T)}u_{ij},
  \label{eq:constructive-tree-graph-inequality}
\end{equation}
where \(\mathcal T_n\) is the set of spanning trees on
\(\{1,\ldots,n\}\).
\end{lemma}

\begin{proof}
Fix an ordering of the edges of the complete graph on \(\{1,\ldots,n\}\).
For a connected graph \(G\), let \(T(G)\) be its lexicographically first
spanning tree.  Penrose's partition scheme assigns to every tree \(T\) a
graph \(R(T)\), containing \(T\), such that the connected graphs are the
disjoint union of intervals
\[
  \{G:\ T\subseteq G\subseteq R(T)\}.
\]
Explicitly, an edge \(e\notin T\) is put in \(R(T)\) precisely when adding
\(e\) to \(T\) does not change the lexicographically first spanning tree.
Therefore
\[
  \sum_{\substack{G\in\mathcal C_n\\T(G)=T}}
  \prod_{e\in E(G)}(-u_e)
  =
  (-1)^{|T|}
  \prod_{e\in E(T)}u_e
  \prod_{e\in E(R(T))\setminus E(T)}(1-u_e).
\]
Since each \(u_e\) is \(0\) or \(1\), the absolute value of the right-hand
side is at most \(\prod_{e\in E(T)}u_e\).  Summing over \(T\) proves
\eqref{eq:constructive-tree-graph-inequality}.
\end{proof}

\begin{theorem}[Kotecky--Preiss/Brydges--Kennedy convergence criterion]
\label{thm:constructive-kp-cluster-convergence}
Let \(\mathcal P_\Lambda\) be the finite set of polymers contained in a
finite cell region \(\Lambda\).  Let
\[
  Z_\Lambda(K)
  =
  \sum_{\mathcal X\subset\mathcal P_\Lambda\ {\rm pairwise\ compatible}}
  \prod_{X\in\mathcal X}K(X)
\]
be the abstract hard-core polymer partition function.  Suppose there are
numbers \(a>a'>0\) and \(\epsilon>0\) such that
\begin{equation}
  \sup_{\Delta_0}
  \sum_{X\ni\Delta_0}
  \kappa(X)e^{a|X|}
  \leq
  \epsilon
  \label{eq:constructive-kp-smallness}
\end{equation}
and \(B_{R_0}\epsilon\leq a'\), where \(B_{R_0}\) is the cell-neighbourhood
constant defined by
\[
  |\{\Delta:\operatorname{dist}_{\rm cell}(\Delta,X)\leq R_0\}|
  \leq B_{R_0}|X|.
\]
Then
\begin{equation}
  \log Z_\Lambda(K)
  =
  \sum_{n\geq1}\frac1{n!}
  \sum_{X_1,\ldots,X_n\in\mathcal P_\Lambda}
  \varphi_{\mathrm c}(X_1,\ldots,X_n)
  \prod_{i=1}^nK(X_i)
  \label{eq:constructive-cluster-expansion}
\end{equation}
is absolutely convergent uniformly in \(\Lambda\).  More precisely, for each
root polymer \(X_0\),
\begin{equation}
  \sum_{n\geq1}\frac1{(n-1)!}
  \sum_{X_2,\ldots,X_n\in\mathcal P_\Lambda}
  |\varphi_{\mathrm c}(X_0,X_2,\ldots,X_n)|
  \prod_{i=2}^n\kappa(X_i)
  \leq
  e^{a'|X_0|}.
  \label{eq:constructive-rooted-tree-bound}
\end{equation}
Consequently connected correlations of observables supported in separated
regions \(A\) and \(B\), whose source activities obey the same local
majorant, satisfy
\begin{equation}
  |\langle F_A G_B\rangle-\langle F_A\rangle\langle G_B\rangle|
  \leq
  C_{F,G}e^{-m_{\mathrm{cl}}\,\operatorname{dist}(A,B)}
  \label{eq:constructive-exponential-clustering-bound}
\end{equation}
for a positive cluster mass \(m_{\mathrm{cl}}\) fixed by the exponential
weight in the norm and by the covariance decay.
\end{theorem}

\begin{proof}
For finite \(\Lambda\), the Mayer identity is a finite algebraic identity.
It expands the logarithm of the hard-core gas in connected incompatibility
graphs and gives \eqref{eq:constructive-cluster-expansion}, with the
coefficient \(\varphi_{\mathrm c}\) defined in
\eqref{eq:constructive-ursell-coefficient}.  Equivalently, it is the formal
Taylor expansion of the analytic branch of \(\log Z_\Lambda\) satisfying
\(\log Z_\Lambda(0)=0\); the absolute convergence proved below is what makes
this formal expansion a uniformly convergent function in the stated
polydisc of activities.

Apply Lemma~\ref{lem:constructive-penrose-tree-graph-bound}.  Root each tree
at the vertex carrying \(X_0\) and orient its edges away from the root.  Let
\(F_N(X)\) be the total majorant of rooted incompatibility trees of depth at
most \(N\), with root polymer \(X\), excluding the activity of the root
itself.  The tree bound gives the recursive domination
\begin{equation}
  F_{N+1}(X)
  \leq
  \exp\!\left(
    \sum_{Y\nsim X}\kappa(Y)F_N(Y)
  \right),
  \qquad F_0(X)=1 .
  \label{eq:constructive-rooted-tree-recursion}
\end{equation}
Assume inductively that \(F_N(Y)\leq e^{a'|Y|}\).  Since \(a>a'\),
\begin{equation}
  \sum_{Y\nsim X}\kappa(Y)e^{a'|Y|}
  \leq
  \sum_{\Delta\in N_{R_0}(X)}
  \sum_{Y\ni\Delta}\kappa(Y)e^{a|Y|}
  \leq
  B_{R_0}\epsilon |X|
  \leq
  a'|X|.
  \label{eq:constructive-kp-recursive-bound}
\end{equation}
Thus \(F_{N+1}(X)\leq e^{a'|X|}\).  Letting \(N\to\infty\) proves
\eqref{eq:constructive-rooted-tree-bound}.  Multiplying by the root
majorant \(\kappa(X_0)\) and summing over \(X_0\) with
\eqref{eq:constructive-kp-smallness} proves absolute convergence of
\eqref{eq:constructive-cluster-expansion}.

For two separated observables, the connected part can only be produced by a
connected cluster intersecting both supports.  Every such cluster contains a
path of incompatible polymers joining the source near \(A\) to the source
near \(B\).  Along such a path,
\begin{equation}
  \sum_{X\ {\rm on\ the\ path}} |X|
  \geq
  c_{R_0}\operatorname{dist}(A,B)
  \label{eq:constructive-path-size-distance}
\end{equation}
for a positive cell-metric constant \(c_{R_0}\).  In the rooted-tree
estimate one may use the \(a'\)-weighted recursion while retaining the
unused factor \(e^{-(a-a')|X|}\) from the \(a\)-weighted smallness norm.
Equation \eqref{eq:constructive-path-size-distance} then gives
\[
  m_{\mathrm{cl}}\geq c_{R_0}(a-a')
\]
up to the additional exponential-decay rate supplied by the massive
Gaussian covariance.  The source activity norms for \(F_A\) and \(G_B\)
supply \(C_{F,G}\).  This proves
\eqref{eq:constructive-exponential-clustering-bound}.
\end{proof}

In the Glimm--Jaffe--Spencer \(P(\phi)_2\) construction, the activities
\(K(X,\phi)\) arise by expanding around cell-local interaction pieces while
retaining large-field regulators.  The theorem above is the abstract
polymer estimate of Kotecky--Preiss and Brydges--Kennedy,\footnote{See
R.~Kotecky and D.~Preiss, ``Cluster expansion for abstract polymer models,''
\emph{Commun. Math. Phys.} 103 (1986), and D.~C. Brydges and T.~Kennedy,
``Mayer expansions and the Hamilton-Jacobi equation,'' \emph{J. Stat. Phys.}
48 (1987).} written here in the hard-core finite-range form needed for the
constructive scalar models.  It is not an optional heuristic: it is the
analytic step that justifies taking \(\Lambda\uparrow\R^2\), proves
exponential clustering in the massive phase, and gives uniform bounds on
the Schwinger functions.

\paragraph{Source derivatives and connected clusters.}
The polymer estimate becomes a QFT statement only after one explains how
local insertions are read from the polymer gas.  Let \(\mathfrak S\) be a
finite set of source labels.  For each \(\alpha\in\mathfrak S\), let
\(A_\alpha\) be the cell neighbourhood of the support of the corresponding
test insertion and let \(b_\alpha(X)\) be a source weight that vanishes unless
\(X\) meets \(A_\alpha\).  At finite volume set
\[
  K_\lambda(X)
  =
  K(X)\exp\!\left(
    \sum_{\alpha\in\mathfrak S}\lambda_\alpha b_\alpha(X)
  \right)
\]
and define \(Z_\Lambda(\lambda)\) by replacing \(K\) with \(K_\lambda\) in the
hard-core partition function.  The connected source coefficient is the finite
identity
\[
  C_{\alpha_1\cdots\alpha_m}^{(\Lambda)}
  =
  \left.
  \partial_{\lambda_{\alpha_1}}\cdots
  \partial_{\lambda_{\alpha_m}}\log Z_\Lambda(\lambda)
  \right|_{\lambda=0}.
\]
Substituting \(K_\lambda\) into the absolutely convergent cluster expansion
gives
\[
\begin{aligned}
  C_{\alpha_1\cdots\alpha_m}^{(\Lambda)}
  &=
  \sum_{n\geq1}\frac1{n!}
  \sum_{X_1,\ldots,X_n\in\mathcal P_\Lambda}
  \varphi_{\rm c}(X_1,\ldots,X_n)
  \prod_{i=1}^nK(X_i)                                      \\
  &\qquad\qquad\times
  \sum_{\tau:\{1,\ldots,m\}\to\{1,\ldots,n\}}
  \prod_{r=1}^m b_{\alpha_r}(X_{\tau(r)}).
\end{aligned}
\]
Thus a nonzero term in a connected \(m\)-point source coefficient contains a
connected incompatibility graph whose vertices collectively meet all source
neighbourhoods.  If two groups of sources lie in cell regions that cannot be
joined by an incompatibility path, the finite partition function factorizes
and the mixed cumulant vanishes.  If they can be joined only through a path of
total polymer size at least proportional to their cell distance, the unused
exponential weight in \eqref{eq:constructive-kp-smallness} gives precisely the
decay used in \eqref{eq:constructive-exponential-clustering-bound}.  For
local fields in \(P(\phi)_2\), the source weights \(b_\alpha\) are obtained by
expanding the regulated source functional together with the local interaction
pieces; the cluster estimate controls the resulting connected Schwinger
distributions after the finite-volume ultraviolet limit has already been
taken.

\begin{quotedtheorem}[Massive \(P(\phi)_2\) OS output from the cluster expansion]
\label{qthm:p-phi-two-cluster-os-output}
Let \(P\) be an even-degree real polynomial bounded below with positive
leading coefficient, let \(m>0\), and choose a reflection-positive
finite-volume ultraviolet regulator.  Assume the finite-volume ultraviolet
limit of Theorem~\ref{thm:p-phi-two-finite-volume-uv-limit} has been taken,
and that the resulting polymer activities satisfy the uniform cluster norm
bound \eqref{eq:constructive-kp-smallness} in the massive phase.  Then the
thermodynamic limits
\[
  S_n(f_1,\ldots,f_n)
  =
  \lim_{\Lambda\uparrow\R^2}S_{\Lambda,n}(f_1,\ldots,f_n)
\]
exist for Schwartz test functions.  The hierarchy \(S=\{S_n\}_{n\geq0}\)
satisfies Euclidean covariance, permutation symmetry, reflection positivity,
distributional regularity, the OS growth hypotheses required for
reconstruction, and exponential clustering of connected Schwinger functions.
\end{quotedtheorem}

The quoted theorem is the constructive \(P(\phi)_2\) output statement.  The
source lineage is the Glimm--Jaffe--Spencer constructive cluster-expansion
program for \(P(\phi)_2\), as presented for example in Glimm--Jaffe's
functional-integral account.  The inputs include the polynomial interaction,
Wick ordering, stability, ultraviolet convergence, the polymer norm estimate,
and preservation of reflection positivity along the regulator sequence.

\paragraph{Spectral consequence of Euclidean clustering.}
The constructive input is the exponential clustering statement in the quoted
theorem; the passage from such a bound to a Hamiltonian gap is a spectral
support calculation.  Assume OS reconstruction has been performed from a scalar
Schwinger hierarchy satisfying Quoted
Theorem~\ref{qthm:p-phi-two-cluster-os-output}.  Let \(A\) be a local
Euclidean-time-zero observable with \(\langle\Omega,A\Omega\rangle=0\), and
suppose its two-point function obeys, for some \(m_{\rm E}>0\),
\[
  0\leq
  \langle\Omega,A^*e^{-\tau H}A\Omega\rangle
  \leq C_A e^{-m_{\rm E}\tau},
  \qquad \tau\geq0 .
\]
Then the spectral measure of \(H\) in the vector \(A\Omega\) is supported in
\([m_{\rm E},\infty)\).  If the set of such local vectors is dense in the
orthogonal complement of the vacuum, the reconstructed Hamiltonian has mass
gap at least \(m_{\rm E}\).

By the spectral theorem,
\[
  \langle\Omega,A^*e^{-\tau H}A\Omega\rangle
  =
  \int_{[0,\infty)}e^{-\tau E}\,\dd\mu_A(E),
  \qquad
  \dd\mu_A(E)=\dd\langle A\Omega,P_H(E)A\Omega\rangle .
\]
If \(\mu_A([0,m_{\rm E}-\delta])>0\) for some \(\delta>0\), then
\[
  \int e^{-\tau E}\,\dd\mu_A(E)
  \geq
  e^{-(m_{\rm E}-\delta)\tau}\mu_A([0,m_{\rm E}-\delta]).
\]
This contradicts the assumed upper bound \(C_Ae^{-m_{\rm E}\tau}\) as
\(\tau\to\infty\).  Therefore no spectral weight lies below \(m_{\rm E}\) in
the vector \(A\Omega\).  If such local vectors are dense in
\(\Omega^\perp\), the spectral projection \(P_H([0,m_{\rm E}))\) annihilates
a dense subspace of \(\Omega^\perp\) and hence vanishes there.  This is the
claimed gap.

\section{Hamiltonian Viewpoint and Number-Operator Bounds}

The original construction of \(\phi^4_2\) can also be formulated in the
Hamiltonian representation.  Let \(\mathcal F\) be the massive one-particle
Fock space on a spatial line or circle, \(H_0\) the free Hamiltonian, and
\(N\) the number operator.  With spatial cutoff \(g\geq0\), the interaction
is
\[
  V(g)=\lambda\int g(x):\!\phi(x)^4\!:\,\dd x .
\]
The decisive estimates are number-operator bounds of the form
\begin{equation}
  \|(N+1)^{-2}V(g)(N+1)^{-2}\|
  \leq C_g|\lambda|,
  \qquad
  \|V(g)\psi\|
  \leq
  \alpha\|H_0\psi\|+\beta\|\psi\|
  \label{eq:constructive-number-operator-bound}
\end{equation}
with \(\alpha<1\) after localization and choice of constants.  These
estimates imply self-adjointness and lower boundedness of
\[
  H(g)=H_0+V(g)+E(g),
\]
where \(E(g)\) is the vacuum-energy coordinate in the regulated Hamiltonian.
The removal of the spatial cutoff is not a formal limit of operators on a
fixed dense domain; it is controlled by spatial cluster estimates and by
convergence of local vacuum expectations.

\begin{remark}[Relation between Hamiltonian and Euclidean constructions]
\label{rem:hamiltonian-euclidean-constructive-equivalence}
The Hamiltonian and Euclidean \(P(\phi)_2\) constructions produce the same
local QFT only after a comparison theorem identifies the Schwinger functions
generated by \(e^{-tH(g)}\) with the Euclidean measure construction and then
passes to the infinite-volume limit.  The common notation \(\phi^4_2\) is not
itself such an identification.
\end{remark}

\section{\texorpdfstring{\(\Phi^4_3\), Phase Cells, and Local Coordinates}{Phi-four-three, Phase Cells, and Local Coordinates}}

In three Euclidean dimensions the quartic scalar theory is
superrenormalizable, but Wick ordering alone does not remove every local
divergence.  A regulator may be given by a momentum cutoff, lattice spacing,
or mollifier.  A typical finite-volume action is
\begin{equation}
  S_{\varepsilon,\Lambda}(\phi)
  =
  \frac12\langle\phi,(-\Delta+m^2)\phi\rangle
  +
  \int_\Lambda
  \left[
    \lambda\phi_\varepsilon^4
    +a_\varepsilon\phi_\varepsilon^2
    +b_\varepsilon
  \right]\dd^3x .
  \label{eq:phi-four-three-regulated-action}
\end{equation}
The coefficients \(a_\varepsilon\) and \(b_\varepsilon\) are local
coordinates of the regulated action.  They are not a second Lagrangian and
are not optional decorations.  Their values are fixed by the renormalization
conditions or by the constructive normalization scheme.

For connected scalar \(\phi^4\) graphs in \(d\) Euclidean dimensions with
\(E\) external legs, \(V\) quartic vertices, and \(L\) loops, the superficial
degree of divergence is
\begin{equation}
  \omega_d
  =
  dL-2I
  =
  d-\frac{d-2}{2}E+(d-4)V .
  \label{eq:phi-four-superficial-degree}
\end{equation}
In \(d=3\), only finitely many local divergences occur: vacuum-energy and
two-point local counterterms.  The four-point and higher connected kernels
are superficially convergent after subdivergences are removed.
The graph identities are
\[
  4V=2I+E,\qquad L=I-V+1 .
\]
Substituting \(I=2V-E/2\) into \(dL-2I\) gives
\[
  d(I-V+1)-2I
  =
  (d-2)I-dV+d
  =
  d-\frac{d-2}{2}E+(d-4)V .
\]
For \(d=3\), this is \(3-E/2-V\).  At fixed \(E\), only finitely many \(V\)
can satisfy \(\omega_3\geq0\), and \(E\geq4\) is convergent except for
subdivergences.  Thus all primitive local divergences are proportional to
the identity or to \(\phi^2\).

With sharp momentum cutoff \(|p|\leq\Lambda\),
\[
  I_d(\Lambda,m)
  =
  \int_{|p|\leq\Lambda}
  \frac{\dd^dp}{(2\pi)^d}\frac{1}{p^2+m^2}
\]
satisfies
\begin{equation}
  I_2(\Lambda,m)
  =
  \frac{1}{2\pi}\log\Lambda+O(1),
  \qquad
  I_3(\Lambda,m)
  =
  \frac{\Lambda}{2\pi^2}+O(1).
  \label{eq:tadpole-cutoff-asymptotics}
\end{equation}
In polar coordinates,
\[
  I_d(\Lambda,m)
  =
  \frac{|S^{d-1}|}{(2\pi)^d}
  \int_0^\Lambda \frac{r^{d-1}}{r^2+m^2}\,\dd r .
\]
For \(d=2\), \(|S^1|=2\pi\) and the radial integral is
\(\frac12\log(\Lambda^2+m^2)-\frac12\log m^2\), giving the coefficient
\((2\pi)^{-1}\).  For \(d=3\), \(|S^2|=4\pi\) and the radial integral is
\(\Lambda-m\arctan(\Lambda/m)\), giving the coefficient \((2\pi^2)^{-1}\).

In the normalization \(\lambda\phi^4\), Wick ordering the quartic interaction
produces
\begin{equation}
  \lambda:\!\phi_\varepsilon^4\!:
  =
  \lambda\phi_\varepsilon^4
  -6\lambda C_\varepsilon(0)\phi_\varepsilon^2
  +3\lambda C_\varepsilon(0)^2.
  \label{eq:phi-four-three-wick-coordinate-shift}
\end{equation}
Thus the leading mass-coordinate divergence is fixed by the six single
contractions in \(\phi^4\).  In three dimensions a logarithmic two-loop
local two-point divergence also appears.  In a BPHZ or phase-cell scheme it
is encoded in the finite-regulator coordinate \(a_\varepsilon\), while the
vacuum diagrams are encoded in \(b_\varepsilon\).

It is useful to separate the Wick-coordinate bookkeeping from the remaining
logarithmic local coordinate.  Write the same cutoff action temporarily in
normal-ordered coordinates as
\begin{equation}
  S_{\varepsilon,\Lambda}
  =
  S_{0,\varepsilon}
  +
  \int_\Lambda
  \left[
    \lambda:\!\phi_\varepsilon^4\!:_\varepsilon
    +\alpha_\varepsilon:\!\phi_\varepsilon^2\!:_\varepsilon
    +\beta_\varepsilon
  \right]\dd^3x .
  \label{eq:phi-four-three-normal-ordered-coordinates}
\end{equation}
Here the subscript on \(:\cdot:_\varepsilon\) records that Wick ordering is
relative to the same cutoff covariance used in the Gaussian reference
measure.  Relative to the un-Wick-ordered coordinates in
\eqref{eq:phi-four-three-regulated-action},
\begin{equation}
  a_\varepsilon
  =
  \alpha_\varepsilon-6\lambda C_\varepsilon(0),
  \qquad
  b_\varepsilon
  =
  \beta_\varepsilon-\alpha_\varepsilon C_\varepsilon(0)
  +3\lambda C_\varepsilon(0)^2 .
  \label{eq:phi-four-three-coordinate-conversion}
\end{equation}
The linear divergence in \(a_\varepsilon\) is therefore the one-loop Wick
conversion; the logarithmic divergence is the leading divergent part of
\(\alpha_\varepsilon\).

This is also the coordinate in which the comparison with stochastic
quantization must be made.  To avoid overloading the symbol \(\lambda\), rename
the Euclidean quartic coefficient in
\eqref{eq:phi-four-three-regulated-action} temporarily as \(g_4\).  If the
same cutoff action is used to generate a Langevin equation in the convention of
Chapter~\ref{ch:constructive-stochastic-quantization-singular-spde}, and if
that equation is written with dynamic cubic coefficient \(\lambda_{\rm dyn}\)
and local linear drift coefficient \(M_\varepsilon\), then
\[
  \lambda_{\rm dyn}=4g_4,\qquad
  M_\varepsilon=-2a_\varepsilon .
\]
Therefore the normal-ordered mass coordinate is
\[
  \alpha_\varepsilon
  =
  -\frac12 M_\varepsilon
  +\frac32\lambda_{\rm dyn}C_\varepsilon(0).
\]
Proposition~\ref{prop:phi-four-three-static-dynamic-mass-chart} expands this
identity in the dynamic BPHZ coordinates.  The comparison is made in
\(\alpha_\varepsilon\), not in the raw drift coefficient, because the one-loop
Wick conversion is a coordinate change between the two presentations.

\begin{proposition}[Vacuum coordinate and normalized Schwinger data]
\label{prop:phi-four-three-vacuum-coordinate-normalization}
Fix a cutoff \(\varepsilon\), a bounded volume \(\Lambda\), and a source
\(J\) supported in \(\Lambda\).  Let \(Z_{\varepsilon,\Lambda}^{(\beta)}[J]\)
be the finite-volume source partition function formed from
\eqref{eq:phi-four-three-normal-ordered-coordinates}.  If
\(\beta_\varepsilon\) is replaced by
\(\beta_\varepsilon+c_\varepsilon\), with
\(\lambda\), \(\alpha_\varepsilon\), and the Gaussian reference measure fixed,
then
\begin{equation}
  Z_{\varepsilon,\Lambda}^{(\beta+c)}[J]
  =
  e^{-c_\varepsilon|\Lambda|}
  Z_{\varepsilon,\Lambda}^{(\beta)}[J].
  \label{eq:phi-four-three-vacuum-coordinate-partition-shift}
\end{equation}
Hence every normalized finite-volume Schwinger function is unchanged:
\begin{equation}
  S_{n,\varepsilon,\Lambda}^{(\beta+c)}(f_1,\ldots,f_n)
  =
  S_{n,\varepsilon,\Lambda}^{(\beta)}(f_1,\ldots,f_n).
  \label{eq:phi-four-three-vacuum-coordinate-normalized-schwinger}
\end{equation}
The connected source functional
\[
  W_{\varepsilon,\Lambda}^{(\beta)}[J]
  :=
  \log Z_{\varepsilon,\Lambda}^{(\beta)}[J]
  -\log Z_{\varepsilon,\Lambda}^{(\beta)}[0]
\]
is also unchanged, while the unnormalized thermodynamic coordinates shift by
\begin{equation}
  p_{\varepsilon,\Lambda}^{(\beta+c)}
  =
  p_{\varepsilon,\Lambda}^{(\beta)}
  -c_\varepsilon,\qquad
  f_{\varepsilon,\Lambda}^{(\beta+c)}
  =
  f_{\varepsilon,\Lambda}^{(\beta)}
  +c_\varepsilon,
  \label{eq:phi-four-three-vacuum-coordinate-pressure-shift}
\end{equation}
where
\(p_{\varepsilon,\Lambda}=|\Lambda|^{-1}\log
Z_{\varepsilon,\Lambda}[0]\) and
\(f_{\varepsilon,\Lambda}=-|\Lambda|^{-1}\log
Z_{\varepsilon,\Lambda}[0]\).
At fixed \(\alpha_\varepsilon\) and \(\lambda\), the same shift changes the
un-Wick-ordered coordinate in
\eqref{eq:phi-four-three-coordinate-conversion} by
\[
  b_\varepsilon\longmapsto b_\varepsilon+c_\varepsilon,
  \qquad
  a_\varepsilon\longmapsto a_\varepsilon .
\]
Thus the stochastic-quantization drift chart determines the mass coordinate
but not the vacuum coordinate; the latter is fixed only after specifying the
partition-function, pressure, or free-energy normalization.
\end{proposition}

\begin{proof}
The replacement
\(\beta_\varepsilon\mapsto\beta_\varepsilon+c_\varepsilon\) adds the constant
\(c_\varepsilon|\Lambda|\) to the finite-volume action.  Pulling this factor
out of the integral gives
\eqref{eq:phi-four-three-vacuum-coordinate-partition-shift}.  Differentiating
the source partition function, or equivalently inserting finitely many
smeared cutoff fields before normalization, multiplies numerator and
denominator by the same factor \(e^{-c_\varepsilon|\Lambda|}\), proving
\eqref{eq:phi-four-three-vacuum-coordinate-normalized-schwinger}.  Taking the
difference
\(\log Z[J]-\log Z[0]\) cancels the same additive logarithmic term, while
dividing \(\log Z[0]-c_\varepsilon|\Lambda|\) by \(|\Lambda|\) gives the
pressure and free-energy shifts in
\eqref{eq:phi-four-three-vacuum-coordinate-pressure-shift}.
Finally, the coordinate conversion
\eqref{eq:phi-four-three-coordinate-conversion} is affine in
\(\beta_\varepsilon\) and independent of \(\beta_\varepsilon\) in
\(a_\varepsilon\).  The functional derivative of an additive constant in the
action is zero, so the Langevin drift sees no \(c_\varepsilon\) term.
\end{proof}

\begin{lemma}[Universal logarithm in the three-dimensional sunset kernel]
\label{lem:phi-four-three-sunset-logarithm}
Let
\[
  C_m(x)=\frac{e^{-m|x|}}{4\pi |x|}
\]
be the massive covariance in \(\R^3\), and let \(C_\varepsilon\) be obtained
from \(C_m\) by a smooth ultraviolet mollifier at length \(\varepsilon\).
For every fixed \(R>0\),
\begin{equation}
  J_\varepsilon(R)
  :=
  \int_{|x|\leq R} C_\varepsilon(x)^3\,\dd^3x
  =
  \frac{1}{16\pi^2}\log\frac1\varepsilon+O_R(1).
  \label{eq:phi-four-three-sunset-log}
\end{equation}
The coefficient of the logarithm is independent of the mollifier.
\end{lemma}

\begin{proof}
Near the origin,
\[
  C_m(x)=\frac{1}{4\pi |x|}+O(1).
\]
On the annulus \(A_{\varepsilon,R}=\{c\varepsilon\leq |x|\leq R\}\), with
\(c\) fixed large enough for the chosen mollifier, standard estimates on
smooth approximate identities give
\[
  C_\varepsilon(x)
  =
  \frac{1}{4\pi |x|}+O(1)+O(\varepsilon |x|^{-2}).
\]
The last error contributes \(O(1)\) to the integral of \(C_\varepsilon^3\)
over \(A_{\varepsilon,R}\).  The ball \(|x|\leq c\varepsilon\) contributes
only a mollifier-dependent bounded term after the logarithmic part has been
extracted, because its volume is \(O(\varepsilon^3)\) and the mollified
covariance is \(O(\varepsilon^{-1})\).  Hence the divergent part is
\[
  \int_{c\varepsilon}^{R}
  4\pi r^2
  \left(\frac{1}{4\pi r}\right)^3\,\dd r
  =
  \frac{1}{16\pi^2}
  \log\frac{R}{c\varepsilon},
\]
which proves \eqref{eq:phi-four-three-sunset-log}.
\end{proof}

\begin{lemma}[Two-loop local mass coordinate in the chapter's normalization]
\label{lem:phi-four-three-two-loop-mass-coordinate}
Use the normal-ordered coordinates
\eqref{eq:phi-four-three-normal-ordered-coordinates}.  Through order
\(\lambda^2\) and first order in \(\alpha_\varepsilon\), the logarithmically
divergent local part of the connected two-point Schwinger function is
\begin{equation}
  \left(96\lambda^2J_\varepsilon(R)-2\alpha_\varepsilon\right)
  \int f(u)C_m(u-z)C_m(z-v)g(v)\,
  \dd^3u\,\dd^3z\,\dd^3v ,
  \label{eq:phi-four-three-two-point-local-divergence}
\end{equation}
up to finite terms and derivative local terms.  Therefore the logarithmic
mass coordinate is
\begin{equation}
  \alpha_\varepsilon
  =
  48\lambda^2J_\varepsilon(R)+\alpha_{\rm fin}+O(\lambda^3)
  =
  \frac{3\lambda^2}{\pi^2}\log\frac1\varepsilon
  +\alpha_{\rm fin}+O(1)+O(\lambda^3),
  \label{eq:phi-four-three-log-mass-coordinate}
\end{equation}
with \(\alpha_{\rm fin}\) fixed by the chosen finite mass normalization.
\end{lemma}

\begin{proof}
The normal-ordered product identity for two quartic vertices is
\begin{equation}
  :\!\phi_\varepsilon(x)^4\!:\,:\!\phi_\varepsilon(y)^4\!:
  =
  \sum_{k=0}^{4}
  \binom{4}{k}^2 k!\,
  C_\varepsilon(x-y)^k
  :\!\phi_\varepsilon(x)^{4-k}
  \phi_\varepsilon(y)^{4-k}\!: .
  \label{eq:phi-four-three-wick-product-two-vertices}
\end{equation}
The logarithmic two-point local term is the \(k=3\) term:
\[
  \binom{4}{3}^2\,3!\,C_\varepsilon(x-y)^3
  :\!\phi_\varepsilon(x)\phi_\varepsilon(y)\!:
  =
  96\,C_\varepsilon(x-y)^3
  :\!\phi_\varepsilon(x)\phi_\varepsilon(y)\!: .
\]
The second order in the expansion of
\(\exp[-\lambda\int:\phi^4:]\) has the factor \(\lambda^2/2\).  Contracting
\(:\!\phi(x)\phi(y)\!:\) with the two external fields gives
\[
  C_m(u-x)C_m(v-y)+C_m(u-y)C_m(v-x).
\]
In the singular region \(x-y\) small, both terms have the same local part.
Thus the quartic vertices contribute
\[
  \frac{\lambda^2}{2}\cdot 96\cdot 2\,J_\varepsilon(R)
  \int f(u)C_m(u-z)C_m(z-v)g(v)\,\dd^3u\,\dd^3z\,\dd^3v ,
\]
which is the \(96\lambda^2J_\varepsilon(R)\) term in
\eqref{eq:phi-four-three-two-point-local-divergence}.  The insertion
\(\alpha_\varepsilon\int:\phi^2:\) contributes at first order
\[
  -\alpha_\varepsilon
  \int \langle\phi(u)\phi(v):\!\phi(z)^2\!:\rangle_0
  \,\dd^3z
  =
  -2\alpha_\varepsilon
  \int C_m(u-z)C_m(z-v)\,\dd^3z .
  \]
Cancellation of the logarithmic coefficient gives
\(\alpha_\varepsilon=48\lambda^2J_\varepsilon(R)+\alpha_{\rm fin}\).  The
coefficient in \eqref{eq:phi-four-three-log-mass-coordinate} follows from
Lemma~\ref{lem:phi-four-three-sunset-logarithm}.  The dependence on the
radius \(R\), on the mollifier, and on the subtraction convention is finite
and is absorbed into \(\alpha_{\rm fin}\).
\end{proof}

\paragraph{Finite-cutoff stability of the local normal-ordered polynomial.}
Fix a cutoff covariance with \(c_\varepsilon=C_\varepsilon(0)<\infty\).
Let \(\lambda>0\), and define the one-site polynomial
\[
  p_\varepsilon(q)
  =
  \lambda:\!q^4\!:_\varepsilon
  +\alpha_\varepsilon:\!q^2\!:_\varepsilon
  +\beta_\varepsilon .
\]
Equivalently,
\[
  p_\varepsilon(q)
  =
  \lambda y^2+A_\varepsilon y+B_\varepsilon,
  \qquad
  y=q^2,\quad
  A_\varepsilon=\alpha_\varepsilon-6\lambda c_\varepsilon,\quad
  B_\varepsilon=3\lambda c_\varepsilon^2-\alpha_\varepsilon c_\varepsilon
  +\beta_\varepsilon .
\]
For all \(q\in\R\),
\begin{equation}
  p_\varepsilon(q)
  \geq
  B_\varepsilon-\frac{(A_\varepsilon^-)^2}{4\lambda},
  \qquad
  A_\varepsilon^-:=\max\{-A_\varepsilon,0\}.
  \label{eq:phi-four-three-local-stability-bound}
\end{equation}
Consequently the finite-cutoff interaction
\[
  \int_\Lambda p_\varepsilon(\phi_\varepsilon(x))\,\dd^3x
\]
is bounded below by
\[
  |\Lambda|\left(
    B_\varepsilon-\frac{(A_\varepsilon^-)^2}{4\lambda}
  \right).
\]
The Wick identities in
Definition~\ref{def:constructive-wick-powers-gaussian-regularization} give
\[
  :\!q^4\!:_\varepsilon=q^4-6c_\varepsilon q^2
  +3c_\varepsilon^2,
  \qquad
  :\!q^2\!:_\varepsilon=q^2-c_\varepsilon .
\]
With \(y=q^2\geq0\), the polynomial becomes
\(\lambda y^2+A_\varepsilon y+B_\varepsilon\).  If
\(A_\varepsilon\geq0\), its minimum on \(y\geq0\) is attained at \(y=0\) and
equals \(B_\varepsilon\).  If \(A_\varepsilon<0\), the unconstrained minimum
lies at \(y=-A_\varepsilon/(2\lambda)>0\) and equals
\(B_\varepsilon-A_\varepsilon^2/(4\lambda)\).  These two cases give
\eqref{eq:phi-four-three-local-stability-bound}.  Integrating the pointwise
bound over \(\Lambda\) proves the final assertion.

This finite-cutoff lower bound does not by
itself prove the continuum \(\Phi^4_3\) theorem, because the lower bound
contains divergent local constants that must be absorbed into
\(\beta_\varepsilon\) and because uniform polymer smallness requires
scale-by-scale estimates on fluctuations rather than only a one-site lower
bound.  It is nevertheless the first stability inequality in the constructive
chain: before phase cells are introduced, the regulated density is an
ordinary integrable function after the local vacuum-energy coordinate is
chosen.

\paragraph{Reflection positivity under split local interactions.}
Let \((\mathfrak A,\theta,\langle\cdot\rangle_0)\) be a reflection-positive
Euclidean probability space, and let \(\mathfrak A_+\subset\mathfrak A\) be
the algebra generated by fields supported at positive Euclidean time.  Thus
\[
  \langle \theta F\,F\rangle_0\geq0,
  \qquad F\in\mathfrak A_+ .
\]
Let \(V_+\) be a real measurable functional affiliated with
\(\mathfrak A_+\), and suppose \(e^{-V_+}F\in L^2\) for the positive-time
test polynomials under consideration.  Define the interacting expectation
by the unnormalized density
\[
  \dd\nu=e^{-V_+-\theta V_+}\,\dd\mu_0 .
\]
Then the unnormalized interacting form is reflection positive:
\begin{equation}
  \langle \theta F\,F\rangle_\nu
  =
  \int \theta F\,F\,e^{-V_+-\theta V_+}\,\dd\mu_0
  \geq0,
  \qquad F\in\mathfrak A_+ .
  \label{eq:constructive-reflection-positivity-split}
\end{equation}

Because \(\theta\) is an anti-linear algebra involution and \(V_+\) is real,
\[
  \theta(e^{-V_+}F)=e^{-\theta V_+}\theta F .
\]
Therefore
\[
  \int \theta F\,F\,e^{-V_+-\theta V_+}\,\dd\mu_0
  =
  \int
  \theta(e^{-V_+}F)\,(e^{-V_+}F)\,\dd\mu_0 .
\]
The right hand side is nonnegative by reflection positivity of the reference
measure, applied to \(e^{-V_+}F\in\mathfrak A_+\).

For a lattice or finite-mode regulator, the split-interaction calculation
above applies directly when the interaction is written as a positive-time
part plus its
reflection.  Terms sitting on the reflection plane are either assigned half
weight before a limiting argument or are treated by a separate
reflection-positive boundary kernel.  Thus reflection positivity in the
constructive theorem is not a formal property of the expression
\(\exp[-S]\); it is a property of the chosen regulator and of the way the
local interaction is split relative to the reflection.

\begin{quotedtheorem}[Constructive \(\Phi^4_3\) output theorem]
\label{thm:phi-four-three-constructive-output}
Let the ultraviolet regulator, finite-volume boundary condition, and mass
coordinate be chosen in one of the massive small-coupling regimes covered by
the constructive \(\Phi^4_3\) phase-cell theorems.  Then there are choices
of local coordinates
\[
  a_\varepsilon=a_{\rm div}(\varepsilon)+a_{\rm fin},\qquad
  b_\varepsilon=b_{\rm div}(\varepsilon)+b_{\rm fin},
\]
such that the regulated Schwinger functions associated with
\eqref{eq:phi-four-three-regulated-action} converge, as
\(\varepsilon\downarrow0\) and \(\Lambda\uparrow\R^3\), to Euclidean
Schwinger distributions satisfying symmetry, Euclidean covariance,
reflection positivity, regularity, clustering in the massive phase, and the
growth hypotheses needed for OS reconstruction.
\end{quotedtheorem}

This is a quoted constructive theorem, not a theorem proved in this
paragraph.  Foundational sources include the Glimm--Jaffe--Spencer
constructive program, Magnen--Seneor's infinite-volume construction of
\(\Phi^4_3\), the Brydges--Federbush phase-cell expansion, and the
presentation in Glimm--Jaffe's \emph{Quantum Physics: A Functional Integral
Point of View}.\footnote{For the constructive lineage relevant here, see
J.~Glimm and A.~Jaffe's constructive-field-theory papers and book,
J.~Magnen and R.~Seneor on the infinite-volume \(\Phi^4_3\) model, and
D.~Brydges and P.~Federbush on phase-cell expansions.  The statement above
is intentionally quoted at theorem-boundary level because reproducing those
multi-scale estimates would require a separate chapter-length proof.}  The
proof is not the formal perturbative cancellation of the diagrams in
\eqref{eq:phi-four-three-wick-coordinate-shift}; the constructive proof has
four analytic components.

In the final standard of this monograph, the quoted theorem is a proof
obligation rather than an authority claim.  The cited constructive literature
identifies the estimates to be checked, but a theorem about the existence of
an interacting QFT should ultimately be accompanied here by a self-contained
derivation of the regulator choice, stability bounds, phase-cell estimates,
uniform infinite-volume convergence, reflection positivity, and OS growth
conditions.

First, the covariance is decomposed into phase cells or scales.  Field
configurations are separated into small-field regions, where perturbative
Taylor expansion is controlled, and large-field regions, where the positive
quartic term and local regulators give stability.

Second, all local divergent pieces allowed by the power-counting formula
\eqref{eq:phi-four-superficial-degree} are extracted into the two
coordinates \(a_\varepsilon\) and \(b_\varepsilon\).  The remaining polymer
activities satisfy bounds of the type
\begin{equation}
  \|K_j\|_{a,b}
  \leq
  C|\lambda|^{1+\delta}L^{-j\alpha}
  \label{eq:phi-four-three-polymer-bound}
\end{equation}
on scale \(j\), for positive \(\delta,\alpha\) in the constructed regime.

Third, the cluster or phase-cell expansion is summed uniformly in the volume
and cutoff.  The convergence criterion is the scale-dependent version of
Theorem~\ref{thm:constructive-kp-cluster-convergence}.  It gives limits of
connected Schwinger functions and exponential clustering.

Fourth, reflection positivity and Euclidean covariance are maintained by the
choice of regulator or recovered in the limiting Schwinger functions.  The
OS growth conditions are checked from uniform moment bounds.  OS
reconstruction then gives the Lorentzian Wightman theory.  Each step is a
theorem with hypotheses on the regulator, mass coordinate, coupling size, and
volume sequence; the Lagrangian label \(\Phi^4_3\) abbreviates this entire
construction.

\begin{definition}[Abstract multiscale phase-cell system]
\label{def:phi-four-three-abstract-phase-cell-system}
Fix a scale factor \(L>1\).  A multiscale phase-cell system consists of:
\begin{enumerate}[label=(\roman*)]
\item cell complexes \(\mathcal D_j\) of side \(L^{-j}\), \(j\geq0\);
\item polymers \(X\subset\mathcal D_j\), with size \(|X|_j\) equal to their
number of \(j\)-cells;
\item a finite-range incompatibility relation \((j,X)\nsim(k,Y)\), whose
cell-neighbourhood constant is \(B_R\);
\item polymer activities \(K_j(X)\) obtained after extracting the local
coordinates \(a_\varepsilon,b_\varepsilon\);
\item numerical majorants \(\kappa_j(X)\geq |K_j(X)|\) after the declared
large-field weights have been applied.
\end{enumerate}
The system has a scale-decay bound with constants
\((C_0,\delta,\alpha,a)\), \(\delta,\alpha,a>0\), if
\begin{equation}
  \sup_{\Delta\in\mathcal D_j}
  \sum_{X\ni\Delta}
  e^{a|X|_j}\kappa_j(X)
  \leq
  C_0|\lambda|^{1+\delta}L^{-\alpha j}
  \quad\hbox{for every }j\geq0 .
  \label{eq:phi-four-three-scale-decay-bound}
\end{equation}
\end{definition}

\begin{theorem}[Summability from the phase-cell scale-decay bound]
\label{thm:phi-four-three-abstract-multiscale-summability}
Assume a multiscale phase-cell system satisfies
\eqref{eq:phi-four-three-scale-decay-bound}.  If \(a>a'>0\) and
\begin{equation}
  B_R\,\frac{C_0|\lambda|^{1+\delta}}{1-L^{-\alpha}}
  \leq a',
  \label{eq:phi-four-three-multiscale-smallness}
\end{equation}
then the combined polymer expansion over all scales is absolutely
convergent.  For a root polymer \((j_0,X_0)\),
\begin{equation}
  \sum_{n\geq1}\frac1{(n-1)!}
  \sum_{(j_2,X_2),\ldots,(j_n,X_n)}
  |\varphi_{\mathrm c}((j_0,X_0),\ldots,(j_n,X_n))|
  \prod_{i=2}^n\kappa_{j_i}(X_i)
  \leq
  e^{a'|X_0|_{j_0}} .
  \label{eq:phi-four-three-multiscale-root-bound}
\end{equation}
Moreover the contribution of scales \(j\geq J\) to the one-root majorant is
bounded by
\begin{equation}
  \frac{C_0|\lambda|^{1+\delta}L^{-\alpha J}}
  {1-L^{-\alpha}} .
  \label{eq:phi-four-three-ultraviolet-tail}
\end{equation}
Therefore the ultraviolet cutoff sequence is Cauchy in every Schwinger
seminorm whose source activities obey the same bound.
\end{theorem}

\begin{proof}
Regard a pair \(\widehat X=(j,X)\) as a single polymer in an enlarged
polymer set.  Its size is \(|\widehat X|:=|X|_j\) and its majorant is
\(\widehat\kappa(\widehat X):=\kappa_j(X)\).  For a fixed cell \(\Delta\)
at any scale, the scale-decay hypothesis gives
\[
  \sum_{j\geq0}\sum_{X\ni\Delta}
  e^{a|X|_j}\kappa_j(X)
  \leq
  \sum_{j\geq0}
  C_0|\lambda|^{1+\delta}L^{-\alpha j}
  =
  \frac{C_0|\lambda|^{1+\delta}}{1-L^{-\alpha}}.
\]
The finite-range constant \(B_R\) gives the neighbourhood estimate
\eqref{eq:constructive-kp-recursive-bound}, with \(\epsilon\) replaced by
this scale sum.  Condition \eqref{eq:phi-four-three-multiscale-smallness}
is the hypothesis needed for the rooted-tree induction in the proof of
Theorem~\ref{thm:constructive-kp-cluster-convergence}.  Applying that proof
to the enlarged polymer set gives
\eqref{eq:phi-four-three-multiscale-root-bound}.

For the ultraviolet tail, the majorant entering the rooted-tree induction is
the restricted scale sum
\[
  \sum_{j\geq J} C_0|\lambda|^{1+\delta}L^{-\alpha j}.
\]
Evaluating this geometric series gives
\[
  \sum_{j\geq J} C_0|\lambda|^{1+\delta}L^{-\alpha j}
  =
  \frac{C_0|\lambda|^{1+\delta}L^{-\alpha J}}
  {1-L^{-\alpha}} .
\]
The rooted-tree recursion is monotone in the majorants.  Thus removing all
scales below \(J\) bounds the remaining ultraviolet tail by
\eqref{eq:phi-four-three-ultraviolet-tail}.  A cutoff change from \(J\) to
\(J'\) changes the Schwinger function only by clusters touching at least one
scale \(j\geq J\), so the same tail estimate proves Cauchy convergence in
the declared source seminorms.
\end{proof}

\begin{construction}[Source-decorated phase-cell seminorms]
\label{con:phi-four-three-source-decorated-phase-cells}
The abstract scale-decay bound becomes a statement about Schwinger
distributions only after local insertions are included in the same norm.
Fix source labels \(1,\ldots,n\), test functions \(f_r\), and seminorms
\(p_r(f_r)\) controlling the corresponding finite-regulator insertion
weights.  A source extension of the phase-cell system consists of functions
\(b_{r,j}(X)\), one for each source label and scale-\(j\) polymer, such that
for every subset \(I\subset\{1,\ldots,n\}\)
\begin{equation}
  \sup_{\Delta\in\mathcal D_j}
  \sum_{X\ni\Delta}
  e^{a|X|_j}\kappa_j(X)
  \prod_{r\in I}|b_{r,j}(X)|
  \leq
  C_I |\lambda|^{1+\delta}L^{-\alpha j}
  \prod_{r\in I}p_r(f_r).
  \label{eq:phi-four-three-source-decorated-scale-bound}
\end{equation}
Here \(I=\varnothing\) recovers
\eqref{eq:phi-four-three-scale-decay-bound}.  At finite volume define
\[
  K_{j,\lambda}(X)
  =
  K_j(X)
  \exp\!\left(\sum_{r=1}^n\lambda_r b_{r,j}(X)\right).
\]
Because the cluster expansion is absolutely convergent in a polydisc
around \(\lambda=0\), finite source derivatives may be taken term by term.
The coefficient
\[
  C_{1\cdots n}^{(J,\Lambda)}
  =
  \left.
  \partial_{\lambda_1}\cdots\partial_{\lambda_n}
  \log Z_{J,\Lambda}(\lambda)
  \right|_{\lambda=0}
\]
is a sum over connected clusters together with an assignment of each source
label to one cluster vertex.  The product over labels assigned to a vertex
is precisely one of the factors controlled in
\eqref{eq:phi-four-three-source-decorated-scale-bound}.  Applying the
tree-graph proof of
Theorem~\ref{thm:phi-four-three-abstract-multiscale-summability} to these
decorated vertices gives constants \(A_n<\infty\), depending on the
neighbourhood constant and on the finite collection \(\{C_I\}_{I\subset
\{1,\ldots,n\}}\), such that
\begin{equation}
  |C_{1\cdots n}^{(J,\Lambda)}|
  \leq
  A_n
  \prod_{r=1}^n p_r(f_r)
  \label{eq:phi-four-three-source-seminorm-bound}
\end{equation}
uniformly in the volume and cutoff.  Moreover, clusters containing at least
one scale \(j\geq J_0\) obey the ultraviolet tail estimate
\begin{equation}
  |C_{1\cdots n}^{(J,\Lambda)}
    -C_{1\cdots n}^{(J_0,\Lambda)}|
  \leq
  A_n
  \frac{|\lambda|^{1+\delta}L^{-\alpha J_0}}
       {1-L^{-\alpha}}
  \prod_{r=1}^n p_r(f_r),
  \qquad J\geq J_0 .
  \label{eq:phi-four-three-source-ultraviolet-cauchy}
\end{equation}
Thus \eqref{eq:phi-four-three-source-decorated-scale-bound} is the missing
source-level hypothesis behind the phrase ``Cauchy in Schwinger seminorms.''
If the constants \(A_n\) can be chosen with an OS-admissible growth envelope,
for instance \(A_n\leq A B^n(n!)^\gamma\) in a nuclear test-space seminorm
family, then the limiting connected Schwinger distributions have the
corresponding OS growth bound.  Proving
\eqref{eq:phi-four-three-source-decorated-scale-bound} and such an
\(n\)-growth envelope for the actual \(\Phi^4_3\) phase-cell activities is
part of the still-open model-specific construction.
\end{construction}

\paragraph{Connected-to-full hierarchy bookkeeping.}
The OS input is the full Schwinger hierarchy, whereas the cluster expansion
first produces connected source coefficients.  This passage is exact but it is
not free at the level of growth constants.  Let \(\Pi_n\) be the set of
partitions of \(\{1,\ldots,n\}\).  If the finite-regulator connected
coefficients \(C_B^{(J,\Lambda)}\), \(B\subset\{1,\ldots,n\}\), are defined by
differentiating \(\log Z_{J,\Lambda}\), then the corresponding full source
coefficient is
\begin{equation}
  S_{1\cdots n}^{(J,\Lambda)}
  =
  \sum_{\pi\in\Pi_n}
  \prod_{B\in\pi} C_B^{(J,\Lambda)} .
  \label{eq:phi-four-three-moment-cumulant-reconstruction}
\end{equation}
This is the finite moment--cumulant identity applied to
\(Z_{J,\Lambda}(\lambda)=\exp\log Z_{J,\Lambda}(\lambda)\).  Taking the
ultraviolet and volume limits term by term is justified only after the
connected coefficients have uniform bounds and Cauchy estimates such as
\eqref{eq:phi-four-three-source-seminorm-bound} and
\eqref{eq:phi-four-three-source-ultraviolet-cauchy}.

The cost in the OS growth envelope is explicit.  Suppose, for one declared
test-seminorm family, that every connected block obeys
\begin{equation}
  |C_B^{(J,\Lambda)}|
  \leq
  A B_0^{|B|}(|B|!)^\gamma
  \prod_{r\in B}p_r(f_r)
  \label{eq:phi-four-three-connected-growth-envelope}
\end{equation}
uniformly in \(J,\Lambda\), with \(A,B_0,\gamma\) independent of the block.
Then \eqref{eq:phi-four-three-moment-cumulant-reconstruction} gives
\begin{equation}
  |S_{1\cdots n}^{(J,\Lambda)}|
  \leq
  B_0^n
  \left(\prod_{r=1}^n p_r(f_r)\right)
  \sum_{\pi\in\Pi_n}
  A^{|\pi|}
  \prod_{B\in\pi}(|B|!)^\gamma .
  \label{eq:phi-four-three-full-from-connected-growth-sum}
\end{equation}
For every partition \(\pi\), \(\prod_{B\in\pi}|B|!\leq n!\).  Also
\(|\Pi_n|\leq n^n\leq e^n n!\).  Hence
\begin{equation}
  |S_{1\cdots n}^{(J,\Lambda)}|
  \leq
  \bigl(e B_0\max\{1,A\}\bigr)^n
  (n!)^{\gamma+1}
  \prod_{r=1}^n p_r(f_r).
  \label{eq:phi-four-three-full-growth-envelope}
\end{equation}
Thus a connected envelope of the OS-II form produces a full hierarchy with the
same linear seminorm-order control but with one additional factorial power from
the partition sum.  This is still admissible for the OS-II condition
\eqref{eq:os-linear-growth-condition} after replacing \(\gamma\) by
\(\gamma+1\), provided the test-seminorm family is stable under the
translation reduction used in Chapter~\ref{ch:os-reconstruction}.  The
statement does not prove the model-specific connected envelope; it records the
exact growth loss that a constructive proof must budget before invoking OS
reconstruction.

The difficult model-specific part of the \(\Phi^4_3\) construction is now
located precisely: one must prove
\eqref{eq:phi-four-three-scale-decay-bound} for the actual phase-cell
activities obtained from the regulated scalar field after the local
coordinates have been extracted.  Theorem
~\ref{thm:phi-four-three-abstract-multiscale-summability} proves that these
bounds, once established with reflection-positive regulators and uniform
moment estimates, are sufficient for ultraviolet convergence of the
Schwinger functions.  Construction
~\ref{con:phi-four-three-source-decorated-phase-cells} records the additional
source-decorated estimate needed to obtain distributional seminorm bounds
and OS growth rather than only convergence of the vacuum polymer partition
function.

\begin{openproblem}[Self-contained constructive proof of \(\Phi^4_3\)]
\label{op:self-contained-phi-four-three-proof}
Give a complete monograph-internal proof of
Quoted Theorem~\ref{thm:phi-four-three-constructive-output}.  The proof must
derive the actual phase-cell activities from the regulated \(\Phi^4_3\)
measure, state the small-field and large-field norms, prove the
model-specific stability estimates beyond the finite-cutoff one-site bound,
identify the local coordinate \(b_\varepsilon\) to the same precision as the
mass coordinate in Lemma~\ref{lem:phi-four-three-two-loop-mass-coordinate},
and prove
\eqref{eq:phi-four-three-scale-decay-bound} and
\eqref{eq:phi-four-three-source-decorated-scale-bound} for those
activities, with OS-admissible growth in the number of source insertions.
The abstract summation, reflection-positivity preservation, and ultraviolet
Cauchy steps are supplied above; the remaining proof obligation is the
model-specific multiscale estimate producing their hypotheses.
\end{openproblem}

\section{OS Data Produced by a Construction}

The output of a constructive scalar theorem is a hierarchy
\begin{equation}
  S_n(f_1,\ldots,f_n)
  =
  \lim_{\Lambda\uparrow\R^d,\ \varepsilon\downarrow0}
  S_{\varepsilon,\Lambda,n}(f_1,\ldots,f_n)
  \label{eq:constructive-schwinger-output}
\end{equation}
on Schwartz test functions or on a declared nuclear test space.  The data
must include the following.

\begin{enumerate}
  \item \textbf{Distributional regularity.}  Each \(S_n\) is a continuous
  multilinear functional in the chosen topology.
  \item \textbf{Euclidean covariance and symmetry.}  The Schwinger functions
  are invariant under the Euclidean group and symmetric under permutation of
  scalar insertions.
  \item \textbf{Reflection positivity.}  For the time reflection
  \(\theta(\tau,\mathbf x)=(-\tau,\mathbf x)\) and test polynomials
  supported at positive Euclidean time,
  \[
    \sum_{\alpha,\beta}
    \overline{c_\alpha}c_\beta
    S_{\alpha+\theta\beta}
    \geq0 .
  \]
  \item \textbf{Growth.}  The OS-II regularity and linear-growth hypotheses
  needed for analytic continuation and Hilbert-space reconstruction are
  verified, not assumed from formal manipulations.
  \item \textbf{Clustering or phase decomposition.}  A unique massive vacuum
  requires exponential clustering; a phase coexistence construction must
  specify the phase sector.
\end{enumerate}

Once these statements are proved, OS reconstruction gives a Hilbert space,
a cyclic vacuum, Wightman fields, and local commutativity.  The constructive
status table in Chapter~\ref{ch:constructive-status-routes} records
representative examples.  The lattice, stochastic, and numerical chapters
below must match the same data format when they claim to approximate or
construct a continuum QFT.

\section{Comparison of Constructive Routes}

The same theory label can arise from cluster expansions, lattice scaling
limits, rigorous RG, and singular SPDE.  Equality of the resulting QFTs is a
statement about Schwinger functions or local observable algebras.  A
comparison theorem must provide a parameter map
\[
  (m,\lambda,a_{\rm fin},b_{\rm fin})_{\rm cluster}
  \longleftrightarrow
  (m,\lambda,c_1,c_2)_{\rm SPDE}
  \longleftrightarrow
  (m_0^2,\lambda_0,a)_{\rm lattice}
\]
and prove convergence to the same \(S_n\)'s in a common topology.

\begin{openproblem}[Explicit comparison of constructive routes]
\label{op:constructive-route-comparison}
For a model such as \(\Phi^4_3\), compare the Schwinger functions produced by
cluster/phase-cell expansions, stochastic quantization, rigorous RG, and
lattice scaling limits in one common topology.  The comparison should
identify the OS Hilbert space, the local fields, and the finite coordinate
map between the constructions.
\end{openproblem}
