\chapter{Stochastic Quantization and Singular SPDE}
\label{ch:constructive-stochastic-quantization-singular-spde}

\ConventionsEuclidean

Stochastic quantization constructs Euclidean QFT data as the invariant law of
a Markov process on fields.  The auxiliary time of the process is not
Lorentzian time and is not an extra physical dimension.  Its purpose is to
replace a formal functional integral by a dynamical construction whose
stationary distribution can sometimes be proved to exist.

In singular dimensions the stochastic equation itself must be renormalized.
The counterterms are local coordinates in the regulated stochastic drift and
must be fixed before the cutoff is removed.  The resulting invariant measure
still has to be checked against the OS hypotheses before it becomes a
Lorentzian QFT by reconstruction.

\section{Finite-Dimensional Langevin Identity}

Let \(S:\R^n\to\R\) be smooth and confining enough that
\[
  Z=\int_{\R^n}\ee^{-S(x)}\,\dd x<\infty .
\]
The overdamped Langevin equation
\begin{equation}
  \dd X_t=-\nabla S(X_t)\,\dd t+\sqrt2\,\dd B_t
  \label{eq:finite-dimensional-langevin}
\end{equation}
has generator
\[
  Lf=-\nabla S\cdot\nabla f+\Delta f .
\]
For smooth compactly supported \(f,g\),
\begin{equation}
  \int (Lf)g\,\ee^{-S}\dd x
  =
  -\int \nabla f\cdot\nabla g\,\ee^{-S}\dd x .
  \label{eq:finite-dimensional-dirichlet-form}
\end{equation}
Therefore \(L\) is symmetric in \(L^2(\ee^{-S}\dd x)\), and the probability
measure \(Z^{-1}\ee^{-S}\dd x\) is invariant.

\begin{proof}
Integrating the drift term by parts gives
\[
  \int (-\partial_iS)(\partial_i f)g\,\ee^{-S}\dd x
  =
  \int \partial_i(\ee^{-S})(\partial_i f)g\,\dd x .
\]
Adding the Laplacian contribution and integrating by parts once more,
\[
  \int
  \bigl[\partial_i(\ee^{-S})(\partial_i f)g
        +(\partial_i^2 f)g\ee^{-S}\bigr]\dd x
  =
  -\int (\partial_i f)(\partial_i g)\ee^{-S}\dd x .
\]
Summing over \(i\) gives \eqref{eq:finite-dimensional-dirichlet-form}.  With
\(g=1\), the right-hand side vanishes, so the adjoint generator annihilates
\(\ee^{-S}\dd x\); this is invariance.
\end{proof}

\section{Regularized Field Equation}

Let \(M\) be a compact Euclidean spatial manifold, let \(C_\Lambda\) be a
smooth ultraviolet cutoff covariance on \(M\), and let \(\xi_\Lambda\) be
white noise in stochastic time with spatial covariance chosen so that the
linear equation has invariant Gaussian law \(\mu_{C_\Lambda}\).  A
regularized scalar \(\Phi^4\) Langevin equation has the form
\begin{equation}
  \partial_\tau\phi_\Lambda
  =
  -C_\Lambda^{-1}\phi_\Lambda
  -\lambda_\Lambda\phi_\Lambda^3
  +c_\Lambda\phi_\Lambda
  +\xi_\Lambda .
  \label{eq:regularized-spde-gradient-flow}
\end{equation}
The drift is the negative \(L^2\)-gradient of the regulated action
\begin{equation}
  S_\Lambda(\phi)
  =
  \frac12(\phi,C_\Lambda^{-1}\phi)
  +
  \int_M
  \left[
    \frac{\lambda_\Lambda}{4}\phi^4
    -\frac{c_\Lambda}{2}\phi^2
  \right]\dd^dx .
  \label{eq:regularized-spde-action}
\end{equation}
At finite \(\Lambda\), the argument of the previous section applies on the
finite-dimensional cutoff field space.  Thus
\[
  Z_\Lambda^{-1}\ee^{-S_\Lambda(\phi)}D\phi
\]
is invariant.  The constructive question is whether the laws of
\(\phi_\Lambda\), with tuned local coordinates, converge as
\(\Lambda\to\infty\).

\section{Renormalized Products}

For the Gaussian free field in spatial dimension \(d\geq2\), \(\phi(x)\) is
a distribution.  Products in the drift are therefore defined through
renormalized local functionals.  If
\(\phi_\epsilon=\rho_\epsilon\ast\phi\) and
\[
  C_\epsilon=\mathbb E[\phi_\epsilon(x)^2],
\]
then
\begin{equation}
  :\!\phi_\epsilon^3\!:
  =
  \phi_\epsilon^3-3C_\epsilon\phi_\epsilon .
  \label{eq:spde-cubic-wick-product}
\end{equation}
The coefficient \(3\) is not conventional: it counts the possible
contractions of two of the three fields.  The same convention gives
\[
  :\!\phi_\epsilon^4\!:
  =
  \phi_\epsilon^4-6C_\epsilon\phi_\epsilon^2+3C_\epsilon^2 .
\]

\section{Da Prato--Debussche in Two Dimensions}

On a two-dimensional torus, consider the stochastic quantization equation
for the Wick-ordered measure
\begin{equation}
  \partial_\tau\Phi
  =
  \Delta\Phi-m^2\Phi-\lambda:\!\Phi^3\!:+\xi .
  \label{eq:dpd-phi-four-two}
\end{equation}
Let \(X\) solve the stationary linear equation
\begin{equation}
  \partial_\tau X=(\Delta-m^2)X+\xi .
  \label{eq:ornstein-uhlenbeck-field}
\end{equation}
The field \(X(\tau,\cdot)\) has the massive Gaussian free-field law on the
spatial torus and is almost surely a spatial distribution of regularity
\(-\kappa\) for every \(\kappa>0\).  Write
\[
  \Phi=X+Y.
\]
Then the nonlinear equation for \(Y\) is
\begin{equation}
  \partial_\tau Y
  =
  (\Delta-m^2)Y
  -\lambda\bigl(
    Y^3+3Y^2X+3Y:\!X^2\!:+:\!X^3\!:
  \bigr).
  \label{eq:dpd-remainder-equation}
\end{equation}
In two dimensions the products in
\eqref{eq:dpd-remainder-equation} are well-defined once
\(:\!X^2\!:\) and \(:\!X^3\!:\) are constructed as Wick powers, because the
unknown \(Y\) gains positive parabolic regularity from the heat operator.

\begin{quotedtheorem}[Da Prato--Debussche solution mechanism]
\label{thm:dpd-solution-mechanism}
Fix \(m^2>0\), \(\lambda>0\), and a finite spatial torus in dimension two.
For almost every realization of the enhanced noise
\[
  \mathbb X=(X,:\!X^2\!:,:\!X^3\!:),
\]
the remainder equation \eqref{eq:dpd-remainder-equation} has local
pathwise solutions in a positive-regularity function space, and the positive
quartic drift supplies a priori bounds that extend the Markov process
globally.  The invariant law is the Wick-ordered \(\Phi^4_2\) Euclidean
measure on the torus.
\end{quotedtheorem}

This Da Prato--Debussche theorem is quoted here as a precise SPDE theorem
boundary, not as a monograph-proved result.  The quoted source is the
Da Prato--Debussche stochastic-quantization construction for strong
solutions.  The mechanism is as follows.
The mild form of \eqref{eq:dpd-remainder-equation} is
\[
  Y(\tau)
  =
  e^{\tau(\Delta-m^2)}Y(0)
  -
  \lambda\int_0^\tau
  e^{(\tau-s)(\Delta-m^2)}
  \mathcal N(Y(s),\mathbb X(s))\,\dd s ,
\]
where \(\mathcal N\) is the cubic expression in
\eqref{eq:dpd-remainder-equation}.  Parabolic Schauder estimates increase
spatial regularity by two derivatives relative to the forcing, while the
roughest forcing \(:\!X^3\!:\) has regularity below zero but above the
threshold needed for the time-integrated heat kernel.  The remaining
products contain at least one factor \(Y\) with positive regularity and are
therefore defined by Besov or Holder multiplication.  These estimates make
the mild map a contraction on a short time interval.  Multiplying the
equation by \(Y\), integrating over space, and using the sign of
\(\lambda Y^4\) gives an energy inequality with random lower-order forcing;
that estimate prevents blowup and gives global solutions.  The invariant law
is identified by applying the finite-cutoff Langevin identity and proving
tightness plus convergence of the enhanced noise.  In the final standard of
this monograph, these Schauder, multiplication, energy, tightness, and
invariant-law identification estimates are proof obligations, not facts
accepted merely because they are standard in the SPDE literature.

\section{\texorpdfstring{Regularity Structures and \(\Phi^4_3\)}{Regularity Structures and Phi-four-three}}

In three spatial dimensions the decomposition \(\Phi=X+Y\) is not
sufficient: \(X\) has regularity \(-\frac12-\kappa\), and the products
generated by the cubic drift fall below the classical multiplication
threshold.  Regularity structures supply the missing local expansion
calculus.

\begin{definition}[Regularity structure]
\label{def:regularity-structure}
A regularity structure is a triple \((A,T,G)\), where \(A\subset\R\) is a
locally finite set of homogeneities bounded below, \(T=\bigoplus_{\alpha\in
A}T_\alpha\) is a graded model space, and \(G\) is a structure group acting
on \(T\) such that for \(\tau\in T_\alpha\),
\[
  \Gamma\tau-\tau\in\bigoplus_{\beta<\alpha}T_\beta
  \qquad(\Gamma\in G).
\]
A model on spacetime assigns to every point \(z\) a linear map
\(\Pi_z:T\to\mathcal D'(\R^{1+d})\) and to every pair \(z,z'\) an element
\(\Gamma_{zz'}\in G\), satisfying
\[
  \Pi_z=\Pi_{z'}\Gamma_{z'z},\qquad
  \Gamma_{zz'}\Gamma_{z'z''}=\Gamma_{zz''},
\]
together with analytic bounds compatible with the homogeneities.
\end{definition}

\begin{quotedtheorem}[Reconstruction theorem, role in SPDE]
\label{thm:regularity-reconstruction-role}
For a model satisfying the analytic bounds of
Definition~\ref{def:regularity-structure}, every modelled distribution
\(F\) of regularity \(\gamma>0\) has a unique distribution \(\mathcal R F\)
such that, near each point \(z\), \(\mathcal R F\) is approximated by
\(\Pi_zF(z)\) to order \(\gamma\).  In singular SPDE, this theorem converts
the formal local expansion in \(T\) into an actual distributional solution.
\end{quotedtheorem}

This is Hairer's reconstruction theorem from the regularity-structures
framework.  It is quoted here as a theorem boundary.  Its role in the
present chapter is the following.  The
modelled-distribution bounds say that the coefficients \(F(z)\) and
\(\Gamma_{zz'}F(z')\) agree up to errors of order
\(|z-z'|^{\gamma-\alpha}\) in each homogeneity \(\alpha\).  Testing the
candidate local distributions \(\Pi_zF(z)\) against wavelets at scale
\(2^{-j}\), these consistency estimates make the wavelet coefficients
Cauchy across neighboring base points.  Summing the coefficients over scales
constructs a distribution.  If two distributions had the same local
approximation bounds, their wavelet coefficients would decay faster than
every scale allowed by the lowest homogeneity, forcing their difference to
vanish.  A self-contained proof in this monograph must spell out the model
norms, wavelet or test-function estimates, summability over scales, and the
uniqueness argument rather than relying on this summary.

For dynamic \(\Phi^4_3\), the symbols are generated by the noise \(\Xi\), the
abstract integration map \(\mathcal I\), multiplication, and polynomials in
the spacetime coordinates.  Negative-homogeneity trees encode precisely the
subdivergent stochastic diagrams.  Renormalization acts on these trees before
the reconstruction theorem is applied.

\section{BPHZ Renormalization of the Dynamic Equation}

For the normalization
\[
  \partial_\tau\Phi
  =
  \Delta\Phi-m^2\Phi-\lambda\Phi^3+\xi ,
\]
the renormalized cutoff equation in the BPHZ regularity-structure convention
has the form
\begin{equation}
  \partial_\tau\Phi_\epsilon
  =
  \Delta\Phi_\epsilon-m^2\Phi_\epsilon
  -\lambda\Phi_\epsilon^3
  +
  \bigl(3\lambda C_{1,\epsilon}
        -9\lambda^2 C_{2,\epsilon}
        +c_{\rm fin}\bigr)\Phi_\epsilon
  +\xi_\epsilon .
  \label{eq:phi-four-three-renormalized-spde}
\end{equation}
Here \(C_{1,\epsilon}\) is the one-loop contraction already visible in
\(:\!\Phi^3\!:\), \(C_{2,\epsilon}\) is the two-loop local contraction
arising from the negative tree with two cubic vertices, and \(c_{\rm fin}\)
is the finite mass coordinate.  Different smooth mollifiers or subtraction
schemes change \(C_{1,\epsilon}\), \(C_{2,\epsilon}\), and \(c_{\rm fin}\)
separately, but not the statement that only the local linear drift and
vacuum-energy coordinates are needed in this superrenormalizable model.

\begin{quotedtheorem}[Renormalized dynamic \(\Phi^4_3\) theorem package]
\label{thm:renormalized-phi-four-three-spde-datum}
Let the spatial manifold be a finite three-dimensional torus, let
\(\xi_\epsilon=\rho_\epsilon\ast\xi\) be a smooth space-time mollification
of white noise, and fix the normalization of
\eqref{eq:phi-four-three-renormalized-spde}.  In the BPHZ
regularity-structure chart there are local constants
\(C_{1,\epsilon}\), \(C_{2,\epsilon}\), and a finite mass coordinate
\(c_{\rm fin}\), such that the following statements hold.
\begin{enumerate}[label=(\roman*)]
  \item For each fixed \(\epsilon>0\) and smooth initial field, the cutoff
  equation \eqref{eq:phi-four-three-renormalized-spde} has a unique smooth
  local solution.  The solution map is Markovian and depends continuously on
  the initial field and on the smooth noise.
  \item After enhancing the mollified noise by its BPHZ-renormalized negative
  trees, the cutoff solutions converge in probability, on compact
  stochastic-time intervals before explosion and in negative spatial Holder
  or Besov topologies, to a limit independent of the mollifier up to finite
  local coordinate changes.
  \item The positive quartic drift, together with the renormalized mass
  coordinate, gives global solutions and invariant probability measures for
  the limiting Markov process in the massive phase.  Starting the cutoff
  equation in its finite-cutoff Gibbs law gives stationary solutions, and
  every limiting stationary law is invariant for the limiting Markov
  semigroup.
  \item In a common finite-volume regulator and renormalization chart, this
  invariant law has the same Schwinger moments as the constructive
  \(\Phi^4_3\) Euclidean measure with the same renormalized mass coordinate.
  Equivalently, after the field and mass normalizations are matched, the
  SPDE stationary construction and the constructive Euclidean construction
  determine the same OS input hierarchy.
\end{enumerate}
\end{quotedtheorem}

This renormalized \(\Phi^4_3\) SPDE theorem is a quoted theorem-boundary for
the regularity-structures construction, not a proof completed here.  The
quoted source lineage is Hairer's regularity-structures construction and the
Hairer--Mourrat--Weber treatments of dynamic \(\Phi^4_3\) renormalization.
The four clauses are logically distinct.  Local cutoff well-posedness is an
ordinary smooth parabolic theorem.  Renormalized convergence is the singular
SPDE theorem: one first builds the canonical model from the mollified noise;
its negative tree components diverge as \(\epsilon\downarrow0\); the BPHZ
character subtracts the expectations of all divergent negative subtrees,
producing a renormalized model satisfying uniform analytic bounds; compactness
and Kolmogorov estimates then give convergence of the model.

The fixed-point problem for the SPDE is then solved in the space of modelled
distributions.  The contraction uses parabolic Schauder estimates for the
abstract integration map and local Lipschitz bounds for the cubic
nonlinearity in the regularity-structure product.  Reconstructing the fixed
point gives the distribution-valued solution.  The explicit local term in
\eqref{eq:phi-four-three-renormalized-spde} is the image, in the concrete
equation, of the BPHZ action on the negative trees.  Invariant measures are
identified by finite-cutoff stationarity, tightness, and convergence of
stationary solutions.  Identification with the constructive Euclidean
\(\Phi^4_3\) measure is still another step: both constructions must be placed
in the same finite-volume cutoff chart, the mass and field coordinates must be
matched, and equality of the limiting Schwinger moments must be proved.  Each
sentence in this paragraph names a theorem-level estimate; none is to be
treated as proved merely by being quoted.

The theorem-boundary statements in this chapter correspond to the
Da Prato--Debussche construction of two-dimensional stochastic
quantization, Hairer's reconstruction theorem, and the regularity-structures
renormalization of dynamic \(\Phi^4_3\).\footnote{See G.~Da Prato and
A.~Debussche, ``Strong solutions to the stochastic quantization equations,''
\emph{Ann. Probab.} 31 (2003); M.~Hairer, ``A theory of regularity
structures,'' \emph{Invent. Math.} 198 (2014); and the
Hairer--Mourrat--Weber regularity-structures treatment of
\(\Phi^4_3\) renormalization.  For recent singular-SPDE work in gauge
contexts, see A.~Chandra, I.~Chevyrev, M.~Hairer, and H.~Shen,
``Stochastic quantisation of Yang--Mills,'' Invent. Math. \textbf{230}
(2022) 1025--1096.}  These references identify the analytic machinery to be
reconstructed in this monograph; they do not by themselves settle the
monograph's proof obligations.

\begin{openproblem}[Self-contained singular-SPDE proof stack]
\label{op:self-contained-singular-spde-proof-stack}
Develop self-contained monograph proofs of the quoted SPDE results used
above.  At minimum this requires: construction and convergence of Wick
powers for the stochastic convolution; the parabolic Schauder and Besov or
Holder multiplication estimates used in the Da Prato--Debussche fixed point;
the global energy estimate and invariant-law identification for
\(\Phi^4_2\); the reconstruction theorem for regularity structures with all
model norms specified; construction of the BPHZ-renormalized model for
dynamic \(\Phi^4_3\); convergence of the renormalized models; solution of
the fixed-point problem in modelled distributions; identification of the
local counterterms in the concrete cutoff equation; and the passage from
stationary SPDE solutions to OS data.  Until these arguments are supplied,
the quoted theorem blocks are documented proof boundaries, not authority
claims.
\end{openproblem}

\section{Paracontrolled and RG Routes}

Regularity structures are not the only analytic language for singular SPDE.
Paracontrolled distributions decompose products using Bony's paraproduct
\[
  fg=f\prec g+f\circ g+f\succ g,
\]
and control the resonant product \(f\circ g\) by enhancing the noise with the
few ill-defined resonant objects needed by the equation.  For \(\Phi^4_3\),
this route uses the same local counterterm content as
\eqref{eq:phi-four-three-renormalized-spde}, expressed in Fourier/Besov
language rather than in a tree Hopf algebra.

Kupiainen-style RG treats the stochastic convolution and nonlinear response
by scale integration.  It is closer in form to the constructive RG of
Chapter~\ref{ch:constructive-rigorous-renormalization-group}.  The common
structure is that no finite truncation of diagrams is being trusted by
itself: the theorem estimates the infinite irrelevant tail in a norm and
shows convergence of the renormalized limit.

\section{Invariant Measure and OS Data}

If the renormalized Markov process exists and is ergodic with invariant
measure \(\mu\), the Euclidean Schwinger functions are
\begin{equation}
  S_n(f_1,\ldots,f_n)
  =
  \int \Phi(f_1)\cdots\Phi(f_n)\,\dd\mu(\Phi).
  \label{eq:spde-invariant-measure-schwinger}
\end{equation}
This is not yet a complete QFT construction.  One must verify:
\[
\begin{gathered}
  \text{Euclidean covariance},\qquad
  \text{reflection positivity},\\
  \text{regularity and OS growth bounds},\qquad
  \text{clustering or phase decomposition}.
\end{gathered}
\]
Stochastic quantization is especially powerful because it gives a dynamic
construction of the measure and a precise algebraic account of stochastic
counterterms.  Reflection positivity, however, is a property of the
Euclidean field measure under physical time reflection; it is not automatic
from Markov positivity in the auxiliary stochastic time.

\begin{remark}[Gauge theories and BV compatibility]
\label{rem:spde-gauge-bv-compatibility}
For gauge theories, stochastic quantization must be reconciled with gauge
symmetry, gauge fixing, and the BV master equation.  A stochastic evolution
for gauge potentials before quotienting is not by itself a construction of a
gauge-invariant local QFT.  The constructive target is a measure or
observable theory for gauge-invariant fields, or a gauge-fixed/BV
construction whose quantum master equation and positivity replacement are
controlled.  Two- and three-dimensional Yang--Mills SPDE programs provide
important laboratories; four-dimensional Yang--Mills remains a separate open
problem.
\end{remark}

\begin{openproblem}[SPDE-to-Wightman construction]
\label{op:spde-to-wightman-construction}
Starting from a renormalized singular SPDE construction of an interacting
Euclidean field, prove the OS hypotheses, including the linear growth
condition needed for reconstruction where applicable, identify the resulting
Wightman distributions, and compare the stochastic counterterms with
Wilsonian local coordinates in a common renormalization chart.
\end{openproblem}
