\chapter{Monte Carlo Methods and Sign Problems}
\label{ch:constructive-monte-carlo-sign-problems}

\ConventionsEuclidean

Monte Carlo methods evaluate finite-regulator Euclidean expectation values by
sampling a probability distribution.  Their mathematical content is a
finite-dimensional integration problem plus a convergence theorem for a
Markov chain.  Their limitation in many QFTs is the absence of a positive
sampling weight compatible with the target observable.

\section{Finite-Regulator Expectation Values}

Let \(\Omega_\Lambda\) be a finite-dimensional regulator space and let
\[
  d\nu_\Lambda(\Phi)=\frac{1}{Z_\Lambda}e^{-S_\Lambda(\Phi)}d\Phi
\]
be a probability measure on it.  For an observable \(O:\Omega_\Lambda\to\mathbb C\),
\[
  \langle O\rangle_\Lambda
  =
  \int_{\Omega_\Lambda}O(\Phi)\,d\nu_\Lambda(\Phi).
\]
On a compact gauge lattice, \(d\Phi\) is a product of Haar measures and
finite-dimensional fermionic integrals have already been performed into a
determinant or Pfaffian when a purely bosonic sampling problem is used.

\section{Markov Chains and Detailed Balance}

A Markov transition kernel \(P(\Phi,d\Phi')\) preserves \(d\nu_\Lambda\) if
\[
  \int_{\Omega_\Lambda}d\nu_\Lambda(\Phi)\,P(\Phi,A)
  =
  \nu_\Lambda(A)
\]
for every measurable set \(A\).  Detailed balance is the stronger condition
\[
  d\nu_\Lambda(\Phi)P(\Phi,d\Phi')
  =
  d\nu_\Lambda(\Phi')P(\Phi',d\Phi).
\]
If the chain is irreducible and aperiodic in the finite-state case, or
satisfies the corresponding Harris recurrence hypotheses in a continuous
state space, the estimator
\[
  \overline O_N=\frac1N\sum_{n=1}^N O(\Phi_n)
\]
converges to \(\langle O\rangle_\Lambda\) in the sense made precise below.

The finite-state statement is elementary enough that it should not be used as
a black box.  Let \(\Omega\) be a finite set, let \(P(x,y)\) be a stochastic
matrix, and let \(\pi(x)>0\) be a probability vector with \(\pi P=\pi\).  A
state \(y\) is reachable from \(x\) if \(P^n(x,y)>0\) for some \(n\).  The
chain is irreducible if every state is reachable from every other state, and
aperiodic if
\[
  \gcd\{n\geq1:P^n(x,x)>0\}=1
\]
for one, hence every, state \(x\).

\begin{proposition}[Finite-state ergodic theorem]
\label{prop:finite-state-markov-ergodic-theorem}
Let \(P\) be an irreducible and aperiodic stochastic matrix on a finite set
\(\Omega\).  Then \(P\) has a unique stationary probability distribution
\(\pi\) with strictly positive entries, and
\begin{equation}
  \lim_{n\to\infty}P^n(x,y)=\pi(y)
  \qquad (x,y\in\Omega).
  \label{eq:finite-chain-pointwise-convergence}
\end{equation}
For every function \(f:\Omega\to\mathbb C\) and every initial state \(X_0\),
\begin{equation}
  \frac1N\sum_{n=0}^{N-1}f(X_n)
  \longrightarrow
  \sum_{x\in\Omega}\pi(x)f(x)
  \label{eq:finite-chain-time-average-convergence}
\end{equation}
in probability.  If the chain is run in stationarity, the same convergence
holds in \(L^2\).
\end{proposition}

\begin{proof}
For an irreducible aperiodic finite chain there is an integer \(m\) for which
\begin{equation}
  P^m(x,y)>0\qquad\text{for all }x,y\in\Omega .
  \label{eq:finite-chain-primitive-power}
\end{equation}
Indeed, irreducibility gives closed walks based at any state, and
aperiodicity makes the semigroup of return times have greatest common divisor
one; adding fixed paths from \(x\) to a reference state and from that
reference state to \(y\) gives positivity for all sufficiently large common
times.  Let
\[
  \epsilon=\min_{x,y}P^m(x,y)>0 .
\]
For probability vectors \(\mu,\nu\), write \(\alpha=\mu-\nu\), so
\(\sum_x\alpha(x)=0\).  Decompose
\[
  P^m(x,y)=\epsilon+R(x,y),
  \qquad
  R(x,y)\geq0,
  \qquad
  \sum_y R(x,y)=1-|\Omega|\epsilon .
\]
The constant part drops out after multiplication by \(\alpha\), and therefore
\begin{equation}
  \|\alpha P^m\|_1
  =
  \|\alpha R\|_1
  \leq
  (1-|\Omega|\epsilon)\|\alpha\|_1 .
  \label{eq:finite-chain-l1-contraction}
\end{equation}
Thus \(P^m\) is a strict contraction on the affine space of probability
vectors modulo total mass.  It has a unique fixed probability vector
\(\pi\), and every \(\mu P^{km}\) converges to \(\pi\).  Since \(\pi P\) is
also fixed by \(P^m\), uniqueness gives \(\pi P=\pi\).  The finitely many
intermediate subsequences \(\mu P^{km+r}\), \(0\leq r<m\), converge to
\(\pi P^r=\pi\), proving \eqref{eq:finite-chain-pointwise-convergence}.  The
positivity of \(\pi\) follows from
\[
  \pi(y)=\sum_x\pi(x)P^m(x,y)\geq\epsilon .
\]

For convergence of time averages in stationarity, it is enough to treat
real-valued \(f\), since the real and imaginary parts can be handled
separately.  Set \(g=f-\sum_x\pi(x)f(x)\).  Since
\(P^n(x,y)\to\pi(y)\), the stationary covariance
\[
  C_g(n)=\sum_{x,y}\pi(x)g(x)P^n(x,y)g(y)
\]
tends to zero.  Hence
\begin{equation}
  \mathbb E_\pi\left|\frac1N\sum_{n=0}^{N-1}g(X_n)\right|^2
  =
  \frac1{N^2}\sum_{r,s=0}^{N-1} C_g(|r-s|)
  \longrightarrow0,
  \label{eq:finite-chain-l2-average-proof}
\end{equation}
because a Cesaro average of a sequence tending to zero tends to zero.  This
gives \(L^2\) convergence in stationarity.  For a nonstationary initial
state, discard a fixed burn-in time \(M\); by
\eqref{eq:finite-chain-pointwise-convergence} the law of \(X_M\) can be made
arbitrarily close to \(\pi\), while the first \(M\) terms have vanishing
weight as \(N\to\infty\).  The convergence in probability follows.
\end{proof}

\section{Metropolis Updates as a Finite-Regulator Theorem}

The simplest exact check of the preceding definitions is a finite spin system.
Let \(\Lambda_L=(\mathbb Z/L\mathbb Z)^2\) be the periodic square lattice and
let a configuration be a map
\[
  s:\Lambda_L\to\{-1,+1\}.
\]
For real \(J,h\), define
\[
  H_L(s)
  =
  -J\sum_{\langle xy\rangle}s_xs_y-h\sum_xs_x ,
  \qquad
  \pi_\beta(s)=Z_{\beta,L}^{-1}e^{-\beta H_L(s)} .
\]
Here \(\langle xy\rangle\) denotes an unoriented nearest-neighbor bond and
\(Z_{\beta,L}\) is the finite sum over all \(2^{L^2}\) configurations.

The single-spin Metropolis chain chooses a site \(x\) uniformly, proposes
\[
  s\longmapsto s^x,\qquad
  s^x_y=
  \begin{cases}
    -s_x,& y=x,\\
    s_y,& y\neq x,
  \end{cases}
\]
and accepts the proposal with probability
\[
  a(s,x)=\min\{1,e^{-\beta(H_L(s^x)-H_L(s))}\}.
\]
The energy difference is local:
\[
  H_L(s^x)-H_L(s)
  =
  2s_x\left(J\sum_{y:\langle xy\rangle}s_y+h\right).
\]

\begin{proposition}[Detailed balance for the finite Ising Metropolis chain]
\label{prop:ising-metropolis-detailed-balance}
The transition kernel \(P\) defined by the preceding update satisfies
\[
  \pi_\beta(s)P(s,s')=\pi_\beta(s')P(s',s)
\]
for all configurations \(s,s'\).  If \(0<\beta<\infty\), the chain is
irreducible.  If additionally \((J,h)\neq(0,0)\), it is aperiodic on the
finite set \(\{-1,+1\}^{\Lambda_L}\).  For the degenerate infinite-temperature
Hamiltonian \(J=h=0\), adding a nonzero holding probability gives the usual
lazy aperiodic variant without changing the stationary measure.
\end{proposition}

\begin{proof}
If \(s'\) is not equal to \(s\) or to a one-spin flip of \(s\), both sides of
the detailed-balance identity vanish.  For \(s'=s^x\), write
\(\Delta H=H_L(s^x)-H_L(s)\).  The off-diagonal transition probabilities are
\[
  P(s,s^x)=\frac1{L^2}\min\{1,e^{-\beta\Delta H}\},
  \qquad
  P(s^x,s)=\frac1{L^2}\min\{1,e^{\beta\Delta H}\}.
\]
If \(\Delta H\geq0\), then
\[
  e^{-\beta H_L(s)}P(s,s^x)
  =
  \frac1{L^2}e^{-\beta H_L(s)}e^{-\beta\Delta H}
  =
  \frac1{L^2}e^{-\beta H_L(s^x)}
  =
  e^{-\beta H_L(s^x)}P(s^x,s).
\]
The case \(\Delta H<0\) is the same identity with \(s\) and \(s^x\)
interchanged.  Dividing by the finite partition function gives detailed
balance.

Irreducibility follows because any configuration can be reached from any
other by flipping the finite set of sites on which they differ, and every
single spin flip has positive proposal probability and positive acceptance
probability.  If \((J,h)\neq(0,0)\), there is a configuration \(s\) and a site
\(x\) with \(H_L(s^x)>H_L(s)\), so the chain has a positive holding
probability at \(s\).  In a finite irreducible Markov chain, one state with a
self-loop forces every state to have period one.  When \(J=h=0\), every
proposal is accepted and the one-spin-flip walk on the hypercube is bipartite;
the standard lazy modification inserts a holding probability and restores
aperiodicity.
\end{proof}

The companion script
\[
  \texttt{qft\_scripts/ising2d\_metropolis.py}
\]
implements exactly this finite chain and reports the acceptance rate, energy
per site, absolute magnetization, and a windowed estimate of
\(\tau_{\mathrm{int}}\).  Its smoke mode is a finite-regulator check of the
algorithmic identities above.  It does not estimate critical exponents, prove
the Onsager scaling limit, or construct the continuum Ising CFT.
The deterministic calculation check
\[
  \texttt{calculation-checks/ising\_metropolis\_finite\_checks.py}
\]
enumerates the \(2\times2\) periodic chain and verifies both the local energy
difference and the detailed-balance identity used by the script.

\section{Single-Link Metropolis for Finite Gauge Fields}

The same finite-regulator theorem should be stated for gauge variables,
because lattice gauge Monte Carlo samples compact group data rather than
coordinate fields.  For the \(\mathbb Z_2\) gauge model of
Proposition~\ref{prop:z2-gauge-character-expansion}, write
\[
  Q(U)=\sum_{p}U_p,\qquad
  \pi_\beta(U)=Z_\Lambda(\beta)^{-1}e^{\beta Q(U)} .
\]
For a link \(e\), let \(U^e\) be the configuration obtained by flipping
\(U_e\).  The only plaquettes whose values change are the plaquettes
containing \(e\).  If
\[
  B_e(U)=\sum_{p\supset e} U_p ,
\]
then
\[
  Q(U^e)-Q(U)=-2B_e(U).
\]
The single-link Metropolis chain chooses a link uniformly and accepts
\(U\mapsto U^e\) with probability
\[
  a(U,e)=\min\{1,\exp(\beta[Q(U^e)-Q(U)])\}.
\]

\begin{proposition}[Detailed balance for the finite \(\mathbb Z_2\) gauge chain]
\label{prop:z2-gauge-metropolis-detailed-balance}
The preceding single-link update satisfies
\[
  \pi_\beta(U)P(U,U')
  =
  \pi_\beta(U')P(U',U)
\]
for all finite lattice gauge configurations \(U,U'\).  On a connected finite
lattice, the chain is irreducible on the full link-configuration space.
Gauge-invariant observables may be averaged either on this full space or
after quotienting by gauge transformations; the full-space chain preserves
the gauge-invariant expectation values because \(\pi_\beta\) is gauge
invariant.
\end{proposition}

\begin{proof}
If \(U'\) differs from \(U\) by more than one link, the off-diagonal
transition probabilities in both directions vanish.  If \(U'=U^e\), set
\(\Delta Q=Q(U^e)-Q(U)\).  The off-diagonal transition probabilities are
\[
  P(U,U^e)=\frac1{|E|}\min\{1,e^{\beta\Delta Q}\},
  \qquad
  P(U^e,U)=\frac1{|E|}\min\{1,e^{-\beta\Delta Q}\}.
\]
For \(\Delta Q\geq0\),
\[
  e^{\beta Q(U)}P(U,U^e)
  =
  \frac1{|E|}e^{\beta Q(U)}
  =
  \frac1{|E|}e^{\beta Q(U^e)}e^{-\beta\Delta Q}
  =
  e^{\beta Q(U^e)}P(U^e,U).
\]
The case \(\Delta Q<0\) is the same identity with \(U\) and \(U^e\)
interchanged.  Division by \(Z_\Lambda(\beta)\) gives detailed balance.

Irreducibility is immediate on the finite link hypercube: any configuration
can be reached from any other by flipping the finite set of links on which
they differ, and each proposed flip has positive acceptance probability for
finite \(\beta\).  Gauge invariance follows because each \(U_p\) and each
closed Wilson loop is unchanged under \(U_{xy}\mapsto g_xU_{xy}g_y\).
Therefore averaging over the full link space includes gauge orbits with equal
weights and gives the same gauge-invariant expectations as any exact
gauge-orbit average with the corresponding orbit weights.
\end{proof}

The companion script
\[
  \texttt{qft\_scripts/z2\_gauge\_3d\_metropolis.py}
\]
implements this update for a periodic three-dimensional cubic lattice.  The
associated deterministic check
\[
  \texttt{calculation-checks/z2\_gauge\_metropolis\_checks.py}
\]
verifies the local score change, pairwise detailed balance, gauge invariance
of the plaquette action and Wilson loops, and the equality between the
\(1\times1\) Wilson-loop average and the plaquette average.

\section{\texorpdfstring{Compact \(SU(2)\) Link Variables}{Compact SU(2) Link Variables}}

The finite \(\mathbb Z_2\) model is useful because it is exactly enumerable
on small lattices, but nonabelian gauge Monte Carlo must also confront the
continuous compact group structure.  On a finite periodic \(4\)-dimensional
hypercubic lattice, assign to each oriented link \(e=(x,\mu)\) a group
element
\[
  U_{x,\mu}\in SU(2),
  \qquad
  U_{x+\hat\mu,-\mu}=U_{x,\mu}^{-1}.
\]
The product Haar measure is finite and normalized.  For an oriented plaquette
\[
  U_{x;\mu\nu}
  =
  U_{x,\mu}U_{x+\hat\mu,\nu}
  U^{-1}_{x+\hat\nu,\mu}U^{-1}_{x,\nu},
\]
define the finite Wilson score
\begin{equation}
  Q(U)
  =
  \sum_{x}\sum_{\mu<\nu}
  \frac12\operatorname{Re}\operatorname{tr}U_{x;\mu\nu}.
  \label{eq:su2-finite-wilson-score}
\end{equation}
The probability measure used by the companion script is
\begin{equation}
  d\nu_{\beta,\Lambda}(U)
  =
  Z_{\beta,\Lambda}^{-1}\exp[\beta Q(U)]
  \prod_{(x,\mu)}dU_{x,\mu},
  \label{eq:su2-finite-wilson-probability}
\end{equation}
where \(dU\) is normalized Haar measure on \(SU(2)\).  This is the finite
probability problem corresponding to the Wilson action convention; continuum
Yang--Mills normalization and the \(a\to0\) scaling trajectory are separate
questions.

A local compact-group Metropolis proposal chooses a link \(e=(x,\mu)\) and an
element \(R\in SU(2)\), and replaces
\[
  U_{x,\mu}\longmapsto R\,U_{x,\mu}.
\]
Let \(\kappa(dR)\) be the proposal law.  The crucial algebraic hypothesis is
inversion symmetry,
\begin{equation}
  \kappa(dR)=\kappa(dR^{-1}).
  \label{eq:su2-proposal-inversion-symmetry}
\end{equation}
For example, representing \(SU(2)\) by unit quaternions
\[
  R=\cos\alpha\,1+\ii\sin\alpha\,\hat n\cdot\sigma,
  \qquad
  \hat n\in S^2,
\]
with an even distribution of \(\alpha\) and a rotation-invariant distribution
of \(\hat n\), satisfies \eqref{eq:su2-proposal-inversion-symmetry}.

\begin{proposition}[Pairwise detailed balance for compact \(SU(2)\) link Metropolis]
\label{prop:su2-link-metropolis-detailed-balance}
For the finite measure \eqref{eq:su2-finite-wilson-probability}, suppose the
link proposal chooses \(e\) uniformly and \(R\) with a law
\(\kappa(dR)=\kappa(dR^{-1})\).  Accept the proposed configuration \(U'\) with
probability
\begin{equation}
  a(U\to U')
  =
  \min\{1,\exp[\beta(Q(U')-Q(U))]\}.
  \label{eq:su2-link-metropolis-acceptance}
\end{equation}
Then the resulting transition kernel satisfies detailed balance with respect
to \(d\nu_{\beta,\Lambda}\).  If the support of \(\kappa\) generates
\(SU(2)\) as a closed subgroup, the chain is irreducible with respect to the
product Haar measure on the compact link space; rejection probabilities give
the usual holding component.
\end{proposition}

\begin{proof}
The left multiplication \(U_e\mapsto R U_e\) preserves Haar measure on the
updated link.  Its inverse move is \(U'_e\mapsto R^{-1}U'_e=U_e\).  Because
the link is chosen with the same probability in both directions and
\(\kappa(dR)=\kappa(dR^{-1})\), the proposal density for the forward move and
the proposal density for the reverse move agree.  It remains only to check
the Metropolis factor.  Set \(\Delta Q=Q(U')-Q(U)\).  If \(\Delta Q\geq0\),
then
\[
  e^{\beta Q(U)} a(U\to U')
  =
  e^{\beta Q(U)}
  =
  e^{\beta Q(U')}e^{-\beta\Delta Q}
  =
  e^{\beta Q(U')} a(U'\to U).
\]
The case \(\Delta Q<0\) is the same identity with \(U\) and \(U'\) exchanged.
Multiplying by the common proposal density and normalizing by
\(Z_{\beta,\Lambda}\) gives detailed balance.  If the support of \(\kappa\)
generates \(SU(2)\), finite products of allowed local left multiplications
can approximate any assignment of link variables by compactness and the group
generation hypothesis.  This gives irreducibility relative to product Haar
measure.  The holding component is present whenever a proposal is rejected;
one may also insert an explicit lazy step without changing the invariant
measure.
\end{proof}

Gauge transformations are maps \(g_x\in SU(2)\) at sites, acting by
\[
  U_{x,\mu}\longmapsto g_xU_{x,\mu}g^{-1}_{x+\hat\mu}.
\]
They leave \eqref{eq:su2-finite-wilson-score} invariant because plaquettes
transform by conjugation:
\[
  U_{x;\mu\nu}\longmapsto g_xU_{x;\mu\nu}g_x^{-1}.
\]
The rectangular Wilson-loop observable in the \((\mu,\nu)\)-plane is
\begin{equation}
  W_{R,T}(x;\mu,\nu)
  =
  \frac12\operatorname{Re}\operatorname{tr}
  \left[
    \prod_{\ell\in \partial \mathcal R_{R,T}(x;\mu,\nu)} U_\ell
  \right],
  \label{eq:su2-rectangular-wilson-loop}
\end{equation}
with inverse links used when the path orientation is opposite to the positive
lattice orientation.  The companion script averages
\eqref{eq:su2-rectangular-wilson-loop} over translations and coordinate
planes.

The script
\[
  \texttt{qft\_scripts/su2\_gauge\_4d\_metropolis.py}
\]
implements the preceding finite Markov chain using the unit-quaternion model
of \(SU(2)\).  The deterministic check
\[
  \texttt{calculation-checks/su2\_gauge\_metropolis\_checks.py}
\]
verifies quaternion group operations, the local plaquette-score change,
pairwise detailed balance, gauge invariance of \(Q(U)\), gauge invariance of
Wilson loops, and the equality between the \(1\times1\) Wilson-loop average
and the average plaquette.  This script is a compact nonabelian
finite-regulator Metropolis demonstration.  Heat-bath, overrelaxation, and
continuum extrapolation require additional regulator data and are not
certified by this example.

\section{Heat-Bath and Overrelaxation Updates}
\label{sec:heat-bath-overrelaxation-finite-regulator}

Heat-bath and overrelaxation updates are most cleanly defined as statements
about conditional probability measures at fixed regulator.  Let
\begin{equation}
  \Omega=X\times Y,\qquad
  d\nu(x,y)=d\nu(x\mid y)\,d\nu_Y(y)
  \label{eq:conditional-factorization}
\end{equation}
be a probability space with regular conditional probability
\(d\nu(x\mid y)\).  A heat-bath update of the \(X\)-coordinate leaves \(y\)
fixed and samples \(x'\) from \(d\nu(x'\mid y)\):
\begin{equation}
  K_{\mathrm{HB}}\bigl((x,y),d x' d y'\bigr)
  =
  d\nu(x'\mid y)\,\delta_y(dy').
  \label{eq:heat-bath-kernel}
\end{equation}

The kernel \eqref{eq:heat-bath-kernel} satisfies detailed balance with
respect to \(d\nu\).  Therefore it preserves \(d\nu\).  A random scan of
finitely many such coordinate heat-bath kernels, with scan probabilities
independent of the present configuration, also preserves \(d\nu\).
The measure on pairs of configurations generated by one heat-bath move is
\begin{equation}
  d\nu_Y(y)\,d\nu(x\mid y)\,d\nu(x'\mid y)\,\delta_y(dy').
  \label{eq:heat-bath-pair-measure}
\end{equation}
This expression is symmetric under \(x\leftrightarrow x'\).  That is exactly
the detailed-balance identity.  A convex linear combination of invariant
kernels is invariant, so a random scan preserves \(d\nu\).

For the finite \(SU(2)\) Wilson measure \eqref{eq:su2-finite-wilson-probability},
fix every link except one link \(U_e\).  The part of the score depending on
\(U_e\) can be written
\begin{equation}
  Q(U)
  =
  \frac12\operatorname{Re}\operatorname{tr}(U_e V_e)
  + Q_{\widehat e}(U_{\widehat e}),
  \label{eq:su2-local-staple-score}
\end{equation}
where \(U_{\widehat e}\) denotes all links other than \(U_e\), and \(V_e\) is
the sum of oriented staples multiplying \(U_e\) in the plaquettes that
contain \(e\).  In the fundamental representation of \(SU(2)\), the set of
matrices
\[
  a_0\,1+\ii\sum_{j=1}^3 a_j\sigma_j,
  \qquad a_j\in\mathbb R,
\]
is closed under addition and multiplication.  Hence, if \(V_e\neq0\), there
are unique \(c_e>0\) and \(W_e\in SU(2)\) such that
\begin{equation}
  V_e=c_e W_e .
  \label{eq:su2-staple-polar-form}
\end{equation}
If \(V_e=0\), the conditional distribution of \(U_e\) is Haar measure.

For \(V_e\neq0\), left or right Haar invariance reduces the conditional
density to the one-parameter form
\begin{equation}
  d\nu(U_e\mid U_{\widehat e})
  =
  Z_e^{-1}
  \exp\!\left[
    \beta c_e\frac12\operatorname{Re}\operatorname{tr}(U_e W_e)
  \right]dU_e .
  \label{eq:su2-link-conditional}
\end{equation}
Writing
\[
  X=U_e W_e=x_0\,1+\ii\,\vec x\cdot\vec\sigma,
  \qquad
  x_0^2+\|\vec x\|^2=1,
\]
the Haar measure on \(SU(2)\simeq S^3\) gives the marginal density
\begin{equation}
  p_{\beta c_e}(x_0)\,dx_0
  =
  \mathcal N_{\beta c_e}^{-1}
  \sqrt{1-x_0^2}\,e^{\beta c_e x_0}\,dx_0,
  \qquad -1\leq x_0\leq1,
  \label{eq:su2-heatbath-x0-density}
\end{equation}
and, conditional on \(x_0\), the direction
\(\vec x/\sqrt{1-x_0^2}\) is uniformly distributed on \(S^2\).  An exact
\(SU(2)\) heat-bath update is precisely an exact sampler of
\eqref{eq:su2-heatbath-x0-density} and this uniform direction, followed by
\begin{equation}
  U_e'=X W_e^{-1}.
  \label{eq:su2-heatbath-reconstruction}
\end{equation}
Different practical heat-bath algorithms are different exact or approximate
ways of sampling \eqref{eq:su2-heatbath-x0-density}; the invariant-measure
statement is the conditional-sampling theorem above.

Overrelaxation is deterministic rather than heat-bath sampling.  At fixed
staple \(W_e\), define
\begin{equation}
  \mathcal O_{W_e}(U_e)
  =
  W_e^{-1}U_e^{-1}W_e^{-1}.
  \label{eq:su2-overrelaxation-map}
\end{equation}
Then
\begin{equation}
  \mathcal O_{W_e}(\mathcal O_{W_e}(U_e))=U_e,
  \qquad
  \frac12\operatorname{Re}\operatorname{tr}
  \bigl(\mathcal O_{W_e}(U_e)W_e\bigr)
  =
  \frac12\operatorname{Re}\operatorname{tr}(U_eW_e).
  \label{eq:su2-overrelaxation-involution-score}
\end{equation}
The first identity says that the map is an involution.  The second follows
because \(\mathcal O_{W_e}(U_e)W_e=(U_eW_e)^{-1}\), and an \(SU(2)\) matrix
has the same real trace as its inverse.  The map preserves Haar measure
because it is a composition of inversion and left/right multiplication by
fixed group elements.

\begin{proposition}[Finite \(SU(2)\) heat-bath and overrelaxation invariance]
\label{prop:su2-heat-bath-overrelaxation-invariance}
At fixed finite lattice and finite \(\beta\), an exact single-link heat-bath
update with conditional density \eqref{eq:su2-link-conditional} satisfies
detailed balance for \eqref{eq:su2-finite-wilson-probability}.  The
single-link overrelaxation map \eqref{eq:su2-overrelaxation-map} preserves
\eqref{eq:su2-finite-wilson-probability}.  If used as a deterministic
involutive kernel, it is reversible with respect to the conditional measure
at fixed staple.  Any finite composition or random mixture of these updates
with Metropolis updates preserves the finite Wilson measure.
\end{proposition}

\begin{proof}
The heat-bath statement is the detailed-balance argument for
\eqref{eq:heat-bath-kernel} applied with \(X=SU(2)\) equal to the selected
link and \(Y\) equal to all other links.  The local
conditional density is \eqref{eq:su2-link-conditional} because every
plaquette containing \(e\) is linear in \(U_e\), and all other plaquettes are
absorbed into \(Q_{\widehat e}\).

For overrelaxation, fix \(U_{\widehat e}\).  Equation
\eqref{eq:su2-overrelaxation-involution-score} says that the local Boltzmann
factor is unchanged, and Haar preservation says that the conditional Haar
measure is unchanged.  Thus the conditional probability measure is preserved.
Because the map is an involution, the measure on pairs
\((U_e,\mathcal O_{W_e}(U_e))\) is symmetric, which is detailed balance for
the deterministic conditional kernel.  Integrating over \(U_{\widehat e}\)
gives preservation of the full finite Wilson measure.  Compositions and
random mixtures of invariant kernels remain invariant.
\end{proof}

Overrelaxation alone is not an ergodic sampling algorithm: it preserves the
local score and is deterministic at fixed staple.  Its role in practice is to
move along approximately constant-action directions between stochastic
updates.  The mathematical invariant-measure statement above is therefore
only one ingredient in a numerical method; it does not by itself establish
mixing, error bars, or continuum scaling.

The companion script
\[
  \texttt{qft\_scripts/su2\_gauge\_4d\_heatbath\_overrelaxation.py}
\]
implements the finite \(SU(2)\) Wilson model with exact single-link
heat-bath conditionals for \eqref{eq:su2-heatbath-x0-density}, interleaved
with a user-chosen number of deterministic overrelaxation sweeps.  It reports
finite-cutoff plaquette and Wilson-loop estimates in JSON format.  Its smoke
mode checks only the finite algorithmic range of the measured observables;
statistical autocorrelation analysis, thermodynamic limits, and continuum
extrapolation require additional runs and estimates.

The deterministic check
\[
  \texttt{calculation-checks/heatbath\_overrelaxation\_checks.py}
\]
verifies the finite conditional heat-bath balance identity, the quaternionic
\(SU(2)\) staple reduction \(V=cW\), the overrelaxation involution, local
score preservation, orthogonality of the overrelaxation map on \(S^3\), and
the local-staple identities used by the companion script.

\section{\texorpdfstring{\(SU(3)\)}{SU(3)} Link Updates and Subgroup Kernels}
\label{sec:su3-link-updates-subgroup-kernels}

Four-dimensional QCD uses \(SU(3)\) rather than \(SU(2)\).  The finite
Wilson measure is still an ordinary compact group integral, but the special
quaternionic simplification used in \eqref{eq:su2-staple-polar-form} no
longer holds.  Let \(U_e\in SU(3)\) be a selected link in a finite pure-gauge
Wilson lattice system and write the local part of the plaquette score as
\begin{equation}
  Q_e(U_e;U_{\widehat e})
  =
  \frac13\operatorname{Re}\operatorname{Tr}_3(U_eV_e)
  +Q_{\widehat e}(U_{\widehat e}).
  \label{eq:su3-local-staple-score}
\end{equation}
Here \(V_e\) is the sum of oriented staples.  It is a complex \(3\times3\)
matrix determined by the other links and is not, in general, a positive scalar
times an element of \(SU(3)\).  Consequently an exact \(SU(3)\) one-link
heat-bath does not reduce to the elementary scalar-density
\eqref{eq:su2-heatbath-x0-density}.

The standard finite-regulator substitute is to use \(SU(2)\) subgroups of
\(SU(3)\).  For \(1\le i<j\le3\), let \(H_{ij}\subset SU(3)\) be the subgroup
which acts as \(SU(2)\) on the span of the \(i\)-th and \(j\)-th color basis
vectors and acts trivially on the orthogonal color line.  Its Lie algebra is
spanned by
\[
  X_{ij}=E_{ij}-E_{ji},\qquad
  Y_{ij}=\ii(E_{ij}+E_{ji}),\qquad
  H_{ij}=\ii(E_{ii}-E_{jj}),
\]
where \(E_{ij}\) is the matrix unit.  These are anti-Hermitian traceless
matrices.

\begin{proposition}[\(SU(2)\) subgroups generate \(SU(3)\)]
\label{prop:su3-generated-by-su2-subgroups}
The connected subgroup of \(SU(3)\) generated by
\[
  H_{12},\qquad H_{13},\qquad H_{23}
\]
is all of \(SU(3)\).
\end{proposition}

\begin{proof}
The six off-diagonal matrices \(X_{ij},Y_{ij}\) for
\((ij)=(12),(13),(23)\) are linearly independent.  The two diagonal matrices
\(H_{12}\) and \(H_{23}\) are linearly independent and span the diagonal
anti-Hermitian traceless subalgebra.  Together these eight matrices form a
basis of \(\mathfrak{su}(3)\).  Therefore the Lie algebra generated by the
three embedded \(\mathfrak{su}(2)\) subalgebras is \(\mathfrak{su}(3)\).  The
subgroup generated by connected subgroups is connected, and the connected
Lie subgroup with Lie algebra \(\mathfrak{su}(3)\) is \(SU(3)\) itself.
\end{proof}

Fix a link \(e\) and one of the subgroups \(H_\alpha\), where
\(\alpha\in\{12,13,23\}\).  A left subgroup update replaces
\[
  U_e\longmapsto hU_e,\qquad h\in H_\alpha,
\]
and leaves all other links fixed.  At fixed \(U_{\widehat e}\), the induced
conditional density on \(H_\alpha\) is
\begin{equation}
  d\nu_{\alpha,e}(h\mid U)
  =
  Z_{\alpha,e}(U)^{-1}
  \exp\!\left[
    \frac{\beta}{3}
    \operatorname{Re}\operatorname{Tr}_3(hU_eV_e)
  \right]dh,
  \label{eq:su3-subgroup-conditional-density}
\end{equation}
where \(dh\) is normalized Haar measure on the embedded \(SU(2)\).

\begin{proposition}[Finite \(SU(3)\) subgroup-update invariance]
\label{prop:su3-subgroup-update-invariance}
At fixed finite lattice and finite \(\beta\), an exact heat-bath draw from
\eqref{eq:su3-subgroup-conditional-density} satisfies detailed balance for the
finite \(SU(3)\) Wilson measure restricted to the selected subgroup orbit.
The same is true for any Metropolis kernel on \(H_\alpha\) whose proposal
measure is invariant under inversion \(h\mapsto h^{-1}\) and whose acceptance
probability is computed from the local score
\eqref{eq:su3-local-staple-score}.  A random scan over finitely many links and
over the three subgroups preserves the finite Wilson measure.
\end{proposition}

\begin{proof}
The heat-bath statement is exactly the detailed-balance argument for
\eqref{eq:heat-bath-kernel} applied to the coordinate \(h\in H_\alpha\) on
the subgroup orbit \(hU_e\).  The conditional
density is \eqref{eq:su3-subgroup-conditional-density} because all plaquettes
not containing \(e\) are absorbed into \(Q_{\widehat e}\), while the plaquettes
containing \(e\) contribute the linear staple term
\(\frac13\operatorname{Re}\operatorname{Tr}_3(hU_eV_e)\).

For the Metropolis statement, let \(r(dh)\) be the proposal law on \(H_\alpha\)
with \(r(A)=r(A^{-1})\).  From a present subgroup coordinate \(h_0\), propose
\(h'h_0\) with \(h'\sim r\), and accept with
\[
  a(h_0,h'h_0)
  =
  \min\{1,\exp[\beta(Q_e(h'h_0U_e)-Q_e(h_0U_e))]\}.
\]
The off-diagonal pairwise detailed-balance identity is the same two-case
identity as in the proof of
Proposition~\ref{prop:ising-metropolis-detailed-balance}, with the symmetric
proposal measure supplied by inversion invariance of \(r\) and Haar measure.
Rejected proposals provide the diagonal holding term.  Finally, finite
compositions and random convex combinations of invariant kernels preserve the
measure.
\end{proof}

\begin{controlledapproximation}[Status of subgroup algorithms]
\label{ca:su3-subgroup-algorithm-status}
Proposition~\ref{prop:su3-subgroup-update-invariance} is an exact
finite-cutoff invariant-measure statement.  It is not a proof of rapid mixing,
of small autocorrelation time, of absence of topological freezing, or of a
continuum extrapolation.  Proposition~\ref{prop:su3-generated-by-su2-subgroups}
explains why subgroup moves have the correct local group-theoretic reach, but
ergodicity of a concrete Markov chain still depends on the proposal support,
the scan schedule, boundary conditions, and the gauge-redundant directions of
the finite lattice system.
\end{controlledapproximation}

The calculation check
\[
  \texttt{calculation-checks/su3\_lattice\_update\_checks.py}
\]
verifies the embedded \(SU(2)\) determinants and unitarity, the
\(\mathfrak{su}(3)\) span, the subgroup commutator normalizations, gauge
covariance of the local staple score, and the finite Metropolis
detailed-balance identity used in the preceding proposition.

The companion data-generator
\[
  \begin{gathered}
  \texttt{qft\_scripts/}\\
  \texttt{su3\_gauge\_4d\_metropolis\_hdf5.py}
  \end{gathered}
\]
implements the finite subgroup-update Metropolis kernel described above for
the periodic four-dimensional pure \(SU(3)\) Wilson lattice model.  Its
normalized plaquette score is
\begin{equation}
  Q(U)=\sum_p\frac13\operatorname{Re}\operatorname{Tr}_3 U_p,
  \qquad
  d\nu_\beta(U)=Z_{\beta,\Lambda}^{-1}e^{\beta Q(U)}
  \prod_{\ell}dU_\ell .
  \label{eq:su3-script-wilson-measure}
\end{equation}
The script writes, when the optional package \(\texttt{h5py}\) is available,
an HDF5 file containing the run parameters, per-sweep plaquette and
rectangular Wilson-loop measurements, the final link field, and the state of
the pseudorandom generator.  This is a checkpoint at the level of the finite
Markov chain: resuming from it continues a regulator computation, not a
continuum construction.  The script can also export the Wilson-loop
measurements as a table with columns
\[
  \texttt{sample},\quad R,\quad T,\quad W ,
\]
which is the sample-level input expected by
\[
  \begin{gathered}
  \texttt{qft\_scripts/}\\
  \texttt{static\_potential\_from\_wilson\_loops.py}.
  \end{gathered}
\]
The deterministic check
\[
  \begin{gathered}
  \texttt{calculation-checks/}\\
  \texttt{su3\_hdf5\_sampler\_checks.py}
  \end{gathered}
\]
verifies the embedded-subgroup proposals, the local score change against a
direct total-score difference, gauge invariance of the score and Wilson loop,
the \(1\times1\) Wilson-loop/plaquette identity, and the HDF5
measurement/checkpoint layout.

The cluster wrapper
\[
  \begin{gathered}
  \texttt{qft\_scripts/cluster/slurm/}\\
  \texttt{su3\_small\_pipeline.sbatch}
  \end{gathered}
\]
runs a small finite-regulator vertical slice: the \(SU(3)\) sampler above,
the Wilson-flow post-processing script of
Chapter~\ref{ch:constructive-wilson-lattice-gauge-theory}, and the
static-potential analysis script of
Chapter~\ref{ch:constructive-wilson-lattice-gauge-theory}.  The surrounding
\(\texttt{qft\_scripts/cluster}\) shell templates record the SSH-control,
source-sync, result-pull, and SLURM-submission protocol.  These files are
execution wrappers only.  They do not change the finite Markov chain, do not
prove mixing, and do not turn the small \(L=2\) smoke job into a continuum
calculation.

The array wrapper
\[
  \begin{gathered}
  \texttt{qft\_scripts/cluster/slurm/}\\
  \texttt{su3\_parameter\_sweep\_array.sbatch}
  \end{gathered}
\]
adds only deterministic parameter bookkeeping.  Its helper
\[
  \texttt{qft\_scripts/cluster/su3\_sweep\_grid.py}
\]
maps a SLURM array index to a finite Cartesian-product coordinate
\((\beta,\texttt{seed})\), writes that coordinate into the task manifest, and
then runs the same finite-regulator sampler, Wilson-flow, and static-potential
pipeline in a task-local directory.  The calculation check
\[
  \texttt{calculation-checks/cluster\_sweep\_grid\_checks.py}
\]
verifies the Cartesian ordering, out-of-range rejection, and shell-assignment
output.  The array index is not a physical or statistical index: it is a
reproducibility coordinate for a family of finite Markov chains.  Mixing,
thermalization, and continuum scaling remain separate mathematical and
statistical claims.

\section{Hybrid Monte Carlo and Pseudofermions}
\label{sec:hmc-pseudofermions-finite-regulator}

Hybrid Monte Carlo is a finite-regulator construction before it is a
continuum algorithm.  The construction begins with a finite-dimensional
configuration manifold \(Q_\Lambda\), a positive reference measure \(dq\),
and a target probability density
\begin{equation}
  d\nu(q)
  =
  Z^{-1}e^{-S(q)}\,dq ,
  \qquad
  Z=\int_{Q_\Lambda}e^{-S(q)}\,dq<\infty .
  \label{eq:hmc-target-measure}
\end{equation}
For a scalar finite lattice, \(Q_\Lambda\) is typically \(\mathbb R^n\) or a
torus.  For a compact gauge lattice, \(Q_\Lambda\) is a finite product of
compact Lie groups and \(dq\) is product Haar measure.  The following theorem
is stated on an abstract finite-dimensional phase space because this is the
part of HMC that is independent of the chosen regulator.

Let \((\Gamma_\Lambda,d\lambda)\) be a finite-dimensional phase space with a
positive reference measure, let \(H:\Gamma_\Lambda\to\mathbb R\) be a
measurable Hamiltonian for which
\begin{equation}
  d\mu_H(z)
  =
  Z_H^{-1}e^{-H(z)}\,d\lambda(z)
  \label{eq:hmc-extended-measure}
\end{equation}
is a probability measure, and let \(R:\Gamma_\Lambda\to\Gamma_\Lambda\) be
an involution preserving \(d\lambda\).  In the usual cotangent-bundle example
\[
  \Gamma_\Lambda=T^*Q_\Lambda,\qquad z=(q,p),
  \qquad R(q,p)=(q,-p).
\]
For \(Q_\Lambda=\mathbb R^n\) with a positive mass matrix \(M\),
\begin{equation}
  H(q,p)=S(q)+\frac12 p^T M^{-1}p,
  \qquad
  d\lambda=dq\,dp .
  \label{eq:hmc-euclidean-hamiltonian}
\end{equation}
The \(q\)-marginal of \(d\mu_H\) is then \eqref{eq:hmc-target-measure} after
normalizing the Gaussian momentum integral.

\begin{definition}[Reversible volume-preserving trajectory]
\label{def:hmc-reversible-volume-preserving}
A numerical trajectory map \(\Phi:\Gamma_\Lambda\to\Gamma_\Lambda\) is
called reversible and volume preserving, with respect to \(R\), if
\begin{equation}
  \Phi_*(d\lambda)=d\lambda,
  \qquad
  R\Phi R=\Phi^{-1}.
  \label{eq:hmc-reversible-volume-preserving}
\end{equation}
The corresponding Metropolis proposal is
\begin{equation}
  G=R\Phi .
  \label{eq:hmc-involutive-proposal}
\end{equation}
\end{definition}

The explicit momentum flip in \(G\) is not a convention with physical
content; it is the algebraic device that makes the proposal an involution.
Indeed \eqref{eq:hmc-reversible-volume-preserving} gives
\[
  G^2=R\Phi R\Phi=\Phi^{-1}\Phi=1 .
\]

\begin{proposition}[HMC Metropolis correction at finite regulator]
\label{prop:hmc-metropolis-correction}
Assume that \(\Phi\) is reversible and volume preserving in the sense of
Definition~\ref{def:hmc-reversible-volume-preserving}.  Let
\begin{equation}
  a(z)
  =
  \min\{1,\exp[-H(Gz)+H(z)]\}
  \label{eq:hmc-acceptance}
\end{equation}
and define the Markov kernel
\begin{equation}
  K(z,B)
  =
  a(z)\mathbf 1_B(Gz)+(1-a(z))\mathbf 1_B(z)
  \label{eq:hmc-metropolis-kernel}
\end{equation}
for measurable \(B\subset\Gamma_\Lambda\).  Then \(K\) satisfies detailed
balance with respect to \(d\mu_H\).  Consequently \(d\mu_H\) is invariant.
\end{proposition}

\begin{proof}
Since \(G\) is an involution and preserves \(d\lambda\), every nontrivial
proposed move is paired with its reverse.  It is enough to prove the
pairwise identity
\begin{equation}
  e^{-H(z)}a(z)=e^{-H(Gz)}a(Gz).
  \label{eq:hmc-pairwise-balance}
\end{equation}
Write \(\Delta H=H(Gz)-H(z)\).  If \(\Delta H\geq0\), then
\[
  e^{-H(z)}a(z)
  =
  e^{-H(z)}e^{-\Delta H}
  =
  e^{-H(Gz)}
  =
  e^{-H(Gz)}a(Gz),
\]
because the reverse move has energy change \(-\Delta H\leq0\) and is
accepted with probability \(1\).  The case \(\Delta H<0\) is the same
calculation with \(z\) and \(Gz\) interchanged.  Multiplying
\eqref{eq:hmc-pairwise-balance} by the common volume element on the paired
points gives detailed balance for the moving part of the kernel.  The
holding part is diagonal and is therefore symmetric with respect to
\(d\mu_H\).  This proves detailed balance for \(K\).
\end{proof}

The usual HMC transition first refreshes \(p\) from its Gaussian conditional
distribution and then applies \eqref{eq:hmc-metropolis-kernel}.  Momentum
refreshment preserves \(d\mu_H\) because it leaves \(q\) fixed and replaces
the momentum by an independent sample from the \(p\)-conditional density.
The composition therefore preserves \(d\mu_H\), and its \(q\)-marginal
preserves \(d\nu\).  Ergodicity, mixing time, acceptance rate, and scaling
toward a continuum limit are separate claims; they do not follow from
detailed balance alone.

\begin{proposition}[Leapfrog reversibility and volume preservation]
\label{prop:leapfrog-reversible-volume-preserving}
Let \(Q_\Lambda=\mathbb R^n\), let \(S\in C^2(\mathbb R^n)\), and let
\(M\) be a positive definite symmetric matrix.  Define the kick and drift
maps
\begin{equation}
  K_a(q,p)=(q,p-a\nabla S(q)),
  \qquad
  D_\epsilon(q,p)=(q+\epsilon M^{-1}p,p),
  \label{eq:hmc-kick-drift}
\end{equation}
and the one-step leapfrog map
\begin{equation}
  L_\epsilon=K_{\epsilon/2}D_\epsilon K_{\epsilon/2}.
  \label{eq:hmc-leapfrog-map}
\end{equation}
Then \(L_\epsilon\) preserves \(dq\,dp\) and satisfies
\begin{equation}
  R L_\epsilon R=L_\epsilon^{-1},
  \qquad R(q,p)=(q,-p).
  \label{eq:hmc-leapfrog-reversibility}
\end{equation}
Every finite power \(L_\epsilon^N\), \(N\in\mathbb Z_{\geq0}\), has the same
two properties.
\end{proposition}

\begin{proof}
The Jacobian matrix of \(K_a\) is block triangular,
\[
  \begin{pmatrix}
    1 & 0\\
    -a\,\operatorname{Hess}S(q) & 1
  \end{pmatrix},
\]
and therefore has determinant \(1\).  The Jacobian matrix of \(D_\epsilon\)
is
\[
  \begin{pmatrix}
    1 & \epsilon M^{-1}\\
    0 & 1
  \end{pmatrix},
\]
also of determinant \(1\).  Hence \(L_\epsilon\) preserves \(dq\,dp\).
Moreover
\[
  RK_aR=K_{-a}=K_a^{-1},
  \qquad
  RD_\epsilon R=D_{-\epsilon}=D_\epsilon^{-1}.
\]
Using the symmetric ordering in \eqref{eq:hmc-leapfrog-map},
\[
  R L_\epsilon R
  =
  K_{-\epsilon/2}D_{-\epsilon}K_{-\epsilon/2}
  =
  L_\epsilon^{-1}.
\]
A finite product of equal leapfrog steps is again reversible in the sense
needed for HMC, because
\[
  R L_\epsilon^N R=(R L_\epsilon R)^N=L_\epsilon^{-N}.
\]
\end{proof}

For a compact gauge field, the same proof is expressed on the cotangent
bundle of a compact Lie group product.  The drift is group multiplication by
the exponential of a Lie-algebra momentum, the kick changes the momentum by
the derivative of the gauge action, and the reference measure is the
Liouville measure built from Haar measure and Lebesgue measure on the Lie
algebra.  The finite-regulator statement remains exactly the same:
volume preservation, reversibility, and the Metropolis correction imply
invariance of the finite target measure.  The construction of globally
smooth forces, the treatment of constraints, and the proof of ergodicity
depend on the chosen lattice action and are additional data.

Pseudofermions enter when the fermion determinant is real and positive after
flavor pairing.  Let \(A\) be a positive Hermitian \(N\times N\) matrix and
write
\[
  d^{2N}\phi=\prod_{j=1}^N d(\operatorname{Re}\phi_j)\,
  d(\operatorname{Im}\phi_j)
\]
for Lebesgue measure on \(\mathbb C^N\).  The elementary identity is the
following finite-dimensional Gaussian integral.

If \(A>0\) is Hermitian, then
\begin{equation}
  \int_{\mathbb C^N}
  \exp[-\phi^\dagger A^{-1}\phi]\,d^{2N}\phi
  =
  \pi^N\det A .
  \label{eq:pseudofermion-determinant-identity}
\end{equation}
Consequently a positive determinant factor \(\det A\) in a finite Euclidean
integral may be represented, up to the \(A\)-independent constant \(\pi^N\),
by the pseudofermion action \(\phi^\dagger A^{-1}\phi\).
Choose a unitary matrix \(U\) such that
\[
  A=U\,\operatorname{diag}(\lambda_1,\ldots,\lambda_N)\,U^\dagger,
  \qquad \lambda_j>0.
\]
The real Jacobian of the unitary change of variables \(\psi=U^\dagger\phi\)
has absolute value \(1\).  Hence the integral factorizes:
\[
  \prod_{j=1}^N
  \int_{\mathbb C}\exp[-|\psi_j|^2/\lambda_j]\,
  d(\operatorname{Re}\psi_j)d(\operatorname{Im}\psi_j)
  =
  \prod_{j=1}^N \pi\lambda_j
  =
  \pi^N\det A.
\]

For a lattice Dirac operator \(D(U)\), two degenerate flavors commonly give
the positive factor
\[
  \det(D^\dagger D)=|\det D|^2,
  \qquad A(U)=D(U)^\dagger D(U)>0
\]
after zero modes have been excluded or regulated.  Equation
\eqref{eq:pseudofermion-determinant-identity} then turns the determinant
into a bosonic Gaussian integral with a nonlocal action involving
\(A(U)^{-1}\).  The nonlocality is not a conceptual defect; it is a finite
linear-algebra statement.  It becomes an algorithmic problem because each
force evaluation requires solving a linear system.

For \(N_f\) degenerate flavors one formally needs
\[
  \det A^{\alpha},
  \qquad \alpha=N_f/2,
\]
and the exact pseudofermion action would contain the matrix function
\(A^{-\alpha}\).  Rational hybrid Monte Carlo replaces \(x^{-\alpha}\) on a
spectral interval \([\lambda_-,\lambda_+]\) containing the spectrum of
\(A(U)\) by a positive rational function \(r_m(x)\).  At finite regulator
this replacement defines the exactly sampled determinant
\begin{equation}
  \pi^{-N}
  \int_{\mathbb C^N}
  \exp[-\phi^\dagger r_m(A)\phi]\,d^{2N}\phi
  =
  \frac{1}{\det r_m(A)}.
  \label{eq:rhmc-rationalized-determinant}
\end{equation}
It equals \(\det A^\alpha\) only if \(r_m(A)=A^{-\alpha}\) on the spectrum,
or if an additional exact correction or reweighting factor is included.
Thus RHMC is not a slogan for ``fractional fermions''; it is an exact
Markov-chain algorithm for a rationalized finite measure, together with a
separate approximation or correction statement comparing that measure to the
desired determinant.

The elementary spectral error estimate used in such a comparison is
pointwise.  If
\begin{equation}
  \|r_m(A)-A^{-\alpha}\|_{\mathrm{op}}\leq\delta ,
  \label{eq:rhmc-spectral-error}
\end{equation}
then for every pseudofermion vector \(\phi\),
\begin{equation}
  \left|
    \phi^\dagger\bigl(r_m(A)-A^{-\alpha}\bigr)\phi
  \right|
  \leq
  \delta\,\|\phi\|^2 .
  \label{eq:rhmc-action-error-bound}
\end{equation}
This bound controls the error in the pseudofermion action for a fixed
configuration and a fixed pseudofermion vector.  Turning it into a uniform
observable bound requires further hypotheses, for example a uniform spectral
gap, a chosen rational approximation, and control of the pseudofermion tail
or an exact reweighting step.  These hypotheses are finite-regulator
algorithmic data, not consequences of QFT axioms.

The deterministic companion check
\[
  \texttt{calculation-checks/hmc\_pseudofermion\_checks.py}
\]
verifies the leapfrog determinant and reversibility identities for a
one-dimensional quadratic action, the pairwise Metropolis balance identity,
the pseudofermion determinant identity after diagonalization, and the
spectral action-error bound \eqref{eq:rhmc-action-error-bound}.

\section{Autocorrelation and Effective Sample Size}

For a stationary chain and a real observable \(O\), define the connected
autocorrelation function
\[
  C_O(t)
  =
  \langle O(\Phi_t)O(\Phi_0)\rangle
  -\langle O\rangle^2 .
\]
The integrated autocorrelation time is
\[
  \tau_{\mathrm{int}}(O)
  =
  \frac12+\sum_{t=1}^\infty\frac{C_O(t)}{C_O(0)}
\]
when the sum converges.  The variance of the sample mean is then
\[
  \operatorname{Var}(\overline O_N)
  =
  \frac{2\tau_{\mathrm{int}}(O)}{N}
  \operatorname{Var}(O)
  +o(N^{-1}).
\]
Critical slowing down is the growth of \(\tau_{\mathrm{int}}\) along a scaling
trajectory toward a continuum limit.

\begin{proposition}[Exact finite-\(N\) variance identity]
\label{prop:markov-autocorrelation-variance-identity}
Assume the chain is stationary and \(O\) is real with \(C_O(0)>0\).  Then
\begin{equation}
  \operatorname{Var}(\overline O_N)
  =
  \frac{1}{N^2}
  \left[
    NC_O(0)+2\sum_{t=1}^{N-1}(N-t)C_O(t)
  \right].
  \label{eq:finite-n-autocorrelation-variance}
\end{equation}
If
\begin{equation}
  \sum_{t=1}^{\infty}|C_O(t)|<\infty ,
  \label{eq:autocorrelation-absolute-summability}
\end{equation}
then
\begin{equation}
  \operatorname{Var}(\overline O_N)
  =
  \frac{2\tau_{\mathrm{int}}(O)}{N}\,C_O(0)+o(N^{-1}).
  \label{eq:asymptotic-autocorrelation-variance}
\end{equation}
\end{proposition}

\begin{proof}
Write \(O_n=O(\Phi_n)\) and
\(\mu_O=\langle O\rangle\).  Stationarity gives
\[
  \operatorname{Cov}(O_r,O_s)=C_O(|r-s|).
\]
Therefore
\[
  \operatorname{Var}(\overline O_N)
  =
  \frac1{N^2}
  \sum_{r,s=0}^{N-1}\operatorname{Cov}(O_r,O_s).
\]
There are \(N\) pairs with \(r=s\).  For each lag \(t\geq1\), there are
\(N-t\) ordered pairs with \(s=r+t\) and \(N-t\) ordered pairs with
\(r=s+t\).  This proves
\eqref{eq:finite-n-autocorrelation-variance}.  Dividing by \(N^{-1}C_O(0)\)
and using absolute summability to replace
\[
  \sum_{t=1}^{N-1}\left(1-\frac{t}{N}\right)\frac{C_O(t)}{C_O(0)}
\]
by its infinite sum gives
\eqref{eq:asymptotic-autocorrelation-variance}.
\end{proof}

\section{Blocking, Jackknife, and Bootstrap Error Coordinates}
\label{sec:blocking-jackknife-bootstrap-error-coordinates}

Autocorrelation estimates enter numerical QFT through finite data-analysis
coordinates.  Let \(O_0,\ldots,O_{N-1}\) be a real time series sampled after
thermalization.  Choose a block length \(b\) and write
\[
  B=\left\lfloor\frac{N}{b}\right\rfloor .
\]
Discarding the final incomplete block for definiteness, define block means
\begin{equation}
  Y_j=\frac1b\sum_{r=0}^{b-1}O_{jb+r},
  \qquad j=0,\ldots,B-1,
  \label{eq:block-mean-definition}
\end{equation}
and
\[
  \overline Y=\frac1B\sum_{j=0}^{B-1}Y_j .
\]
If \(b\) is large compared with the integrated autocorrelation scale, the
block means are closer to independent samples than the original time series.
This sentence is a diagnostic hypothesis about the Markov chain and the
chosen observable; it is checked by varying \(b\), by monitoring slow modes,
and by comparing independent chains.

\begin{remark}[Delete-one-block jackknife for the mean]
\label{rem:block-jackknife-mean-identity}
Assume \(B\ge2\).  Let
\begin{equation}
  s_Y^2=\frac{1}{B-1}\sum_{j=0}^{B-1}(Y_j-\overline Y)^2 .
  \label{eq:block-sample-variance}
\end{equation}
The blocked standard-error coordinate for \(\overline Y\) is
\begin{equation}
  \operatorname{se}_{\rm block}(\overline Y)=\sqrt{\frac{s_Y^2}{B}}.
  \label{eq:block-standard-error}
\end{equation}
Let \(\overline Y_{(j)}\) be the mean of the \(B-1\) block means with the
\(j\)-th block removed:
\begin{equation}
  \overline Y_{(j)}
  =
  \frac{1}{B-1}\sum_{k\ne j}Y_k .
  \label{eq:delete-one-block-mean}
\end{equation}
Then the delete-one-block jackknife variance for the mean,
\begin{equation}
  \operatorname{Var}_{\rm jack}(\overline Y)
  =
  \frac{B-1}{B}\sum_{j=0}^{B-1}
  \left(
    \overline Y_{(j)}
    -
    \frac1B\sum_{k=0}^{B-1}\overline Y_{(k)}
  \right)^2,
  \label{eq:delete-one-block-jackknife-variance}
\end{equation}
equals \(s_Y^2/B\).
Since
\[
  \overline Y_{(j)}
  =
  \frac{B\overline Y-Y_j}{B-1},
  \]
the average of the delete-one estimates is
\[
  \frac1B\sum_{j=0}^{B-1}\overline Y_{(j)}=\overline Y .
  \]
Therefore
\[
  \overline Y_{(j)}-\overline Y
  =
  -\frac{Y_j-\overline Y}{B-1}.
  \]
Substitution in
\eqref{eq:delete-one-block-jackknife-variance} gives
\[
  \operatorname{Var}_{\rm jack}(\overline Y)
  =
  \frac{B-1}{B}
  \frac{1}{(B-1)^2}
  \sum_{j=0}^{B-1}(Y_j-\overline Y)^2
  =
  \frac{s_Y^2}{B}.
  \]
\end{remark}

The preceding remark is the scalar case of the error-coordinate needed
for Wilson-loop ratios, flowed-energy couplings, and hadron effective masses.
Let \(Y_j\in\mathbb R^m\) be block means for \(m\) elementary observables,
let
\begin{equation}
  \overline Y=\frac1B\sum_{j=0}^{B-1}Y_j,
  \qquad
  d_j=Y_j-\overline Y,
  \qquad
  S_Y=\frac{1}{B-1}\sum_{j=0}^{B-1}d_jd_j^{\mathsf T},
  \label{eq:block-vector-covariance}
\end{equation}
and let
\begin{equation}
  \overline Y_{(j)}
  =
  \frac{B\overline Y-Y_j}{B-1}.
  \label{eq:vector-delete-one-block-mean}
\end{equation}
For a smooth scalar analysis function \(F:U\to\mathbb R\), where
\(U\subset\mathbb R^m\) is open and contains the convex hull of the
\(\overline Y_{(j)}\), define
\[
  F_{(j)}=F(\overline Y_{(j)}),
  \qquad
  \widehat F_\bullet=\frac1B\sum_{j=0}^{B-1}F_{(j)},
\]
and
\begin{equation}
  \widehat V_{\rm jack}(F)
  =
  \frac{B-1}{B}\sum_{j=0}^{B-1}(F_{(j)}-\widehat F_\bullet)^2 .
  \label{eq:nonlinear-block-jackknife-variance}
\end{equation}

\begin{proposition}[Correlated delete-one-block jackknife for smooth ratios]
\label{prop:correlated-delete-one-block-jackknife}
Assume \(B\ge2\).  Let \(g=DF_{\overline Y}\) be the differential of \(F\) at
\(\overline Y\), regarded as a row vector.  Suppose that the operator norm of
the Hessian satisfies
\[
  \|D^2F_x\|_{\rm op}\le M
\]
on the convex hull of the delete-one means.  Put
\[
  D_2=\left(\sum_{j=0}^{B-1}\|d_j\|^2\right)^{1/2},
  \qquad
  D_4=\left(\sum_{j=0}^{B-1}\|d_j\|^4\right)^{1/2},
\]
and
\[
  A=\frac{\|g\|D_2}{B-1},
  \qquad
  E=\frac{M D_4}{(B-1)^2}.
\]
Then
\begin{equation}
  \left|
    \widehat V_{\rm jack}(F)
    -
    \frac1B\,gS_Yg^{\mathsf T}
  \right|
  \le
  \frac{B-1}{B}(2AE+E^2).
  \label{eq:nonlinear-jackknife-taylor-bound}
\end{equation}
\end{proposition}

\begin{proof}
Taylor's theorem gives
\[
  F(\overline Y_{(j)})
  =
  F(\overline Y)+g(\overline Y_{(j)}-\overline Y)+R_j,
  \qquad
  |R_j|\le\frac{M}{2}\|\overline Y_{(j)}-\overline Y\|^2 .
\]
Since
\[
  \overline Y_{(j)}-\overline Y=-\frac{d_j}{B-1}
\]
and \(\sum_j d_j=0\), the average of the linear terms is zero.  Therefore the
linear part of \(\widehat V_{\rm jack}(F)\) is
\[
  \frac{B-1}{B}
  \sum_{j=0}^{B-1}
  \left(\frac{gd_j}{B-1}\right)^2
  =
  \frac1B\,gS_Yg^{\mathsf T}.
\]
Let
\[
  a_j=-\frac{gd_j}{B-1},
  \qquad
  e_j=R_j-\frac1B\sum_{k=0}^{B-1}R_k .
\]
Then the difference between the full jackknife variance and its linear part is
\[
  \frac{B-1}{B}
  \left(
    2\sum_{j=0}^{B-1}a_je_j+\sum_{j=0}^{B-1}e_j^2
  \right).
\]
Cauchy's inequality bounds
\[
  \left(\sum_j a_j^2\right)^{1/2}\le A.
\]
Also
\[
  \left(\sum_j e_j^2\right)^{1/2}
  \le
  2\left(\sum_j R_j^2\right)^{1/2}
  \le E .
\]
Substituting these two bounds proves
\eqref{eq:nonlinear-jackknife-taylor-bound}.
\end{proof}

For the static potential, the elementary block vector contains the positive
Wilson-loop coordinates \(W(n,m)\) needed for a chosen finite list of
rectangles.  The analysis functions are
\begin{equation}
  F^{\rm eff}_{n,m}(W)
  =
  -a^{-1}\log\frac{W(n,m+1)}{W(n,m)}
  \label{eq:effective-mass-analysis-function}
\end{equation}
and
\begin{equation}
  F^\chi_{n,m}(W)
  =
  -\log
  \frac{W(n,m)W(n-1,m-1)}
       {W(n,m-1)W(n-1,m)} .
  \label{eq:creutz-analysis-function}
\end{equation}
They are smooth on the positive orthant.  Applying the jackknife to
\(F^{\rm eff}_{n,m}\) or \(F^\chi_{n,m}\) after deleting an entire Monte Carlo
block is therefore a correlated calculation: the covariance among the
rectangles is the matrix \(S_Y\) in
\eqref{eq:block-vector-covariance}.  Propagating the four Wilson-loop errors
in a Creutz ratio independently is a different approximation, justified only
after a separate argument that the relevant off-diagonal covariance entries
are negligible in the chosen data-analysis scheme.

Block bootstrap resampling similarly resamples the block means or the original
contiguous blocks and recomputes the full nonlinear analysis function on every
resampled data set.  Its consistency requires an asymptotic regime in which
\(b\to\infty\), \(B\to\infty\), and the block length grows slowly enough
compared with the total chain length; the finite script described below
reports the coordinate values and leaves the asymptotic claim to the stated
Markov-chain hypotheses.

The companion script
\[
  \texttt{qft\_scripts/autocorrelation\_resampling.py}
\]
computes the biased autocorrelation estimate, a windowed
\(\tau_{\rm int}\), blocked standard errors, delete-one-block jackknife
errors for the mean, and block-bootstrap errors from a one-column CSV time
series.  It is intended as the upstream error-analysis layer for Wilson-loop
and flowed-observable measurements before ratios such as
\eqref{eq:lattice-static-potential-effective-mass} are formed.

\section{Reweighting and the Sign Problem}

Suppose the target finite-regulator integral has complex weight
\[
  e^{-S(\Phi)}=e^{-S_R(\Phi)}e^{\ii\Theta(\Phi)}
\]
with positive reference measure
\[
  d\nu_R=\frac{1}{Z_R}e^{-S_R(\Phi)}d\Phi .
\]
Then
\[
  \langle O\rangle
  =
  \frac{\langle Oe^{\ii\Theta}\rangle_R}
  {\langle e^{\ii\Theta}\rangle_R}.
\]
The denominator is the average phase.  If the free-energy densities of the
target and phase-quenched systems differ by \(\Delta f\) in volume \(V\), then
\[
  \left|\langle e^{\ii\Theta}\rangle_R\right|
  =
  \exp[-V\Delta f].
\]
Here
\[
  Z=\int e^{-S_R(\Phi)}e^{\ii\Theta(\Phi)}\,d\Phi,\qquad
  Z_R=\int e^{-S_R(\Phi)}\,d\Phi,
  \qquad
  \langle e^{\ii\Theta}\rangle_R=\frac{Z}{Z_R},
\]
and \(\Delta f\) is the real number defined by
\begin{equation}
  \Delta f=-\frac1V\log\left|\frac{Z}{Z_R}\right|.
  \label{eq:phase-quenched-free-energy-difference}
\end{equation}
Estimating a ratio with a denominator of this size requires a number of
independent samples growing as \(\exp[2V\Delta f]\) for fixed relative error,
unless the numerator and denominator are computed by a method that avoids
phase reweighting.

\begin{proposition}[Finite-volume phase-reweighting variance bound]
\label{prop:finite-volume-sign-problem-variance}
Let
\[
  m=\langle e^{\ii\Theta}\rangle_R\neq0 .
\]
For \(N\) independent samples from \(d\nu_R\), the unbiased estimator
\[
  \widehat m_N=\frac1N\sum_{j=1}^{N}e^{\ii\Theta(\Phi_j)}
\]
satisfies
\begin{equation}
  \frac{\mathbb E_R|\widehat m_N-m|^2}{|m|^2}
  =
  \frac{1-|m|^2}{N|m|^2}.
  \label{eq:average-phase-relative-variance}
\end{equation}
Consequently, if \(|m|=\exp[-V\Delta f]\), any direct phase-reweighting
estimation of the denominator with relative mean-square error at most
\(\delta^2\) requires
\begin{equation}
  N\geq
  \delta^{-2}\left(e^{2V\Delta f}-1\right).
  \label{eq:sign-problem-sample-lower-bound}
\end{equation}
\end{proposition}

\begin{proof}
Since \(|e^{\ii\Theta}|=1\),
\[
  \mathbb E_R|e^{\ii\Theta}-m|^2
  =
  \mathbb E_R|e^{\ii\Theta}|^2-|m|^2
  =
  1-|m|^2 .
  \]
Independence divides this variance by \(N\), proving
\eqref{eq:average-phase-relative-variance}.  Substitution of
\(|m|=e^{-V\Delta f}\) gives
\eqref{eq:sign-problem-sample-lower-bound}.  The estimate concerns the
denominator alone.  A ratio estimator for \(\langle O\rangle\) can have
additional covariance between numerator and denominator, but it cannot remove
the need to determine a denominator of exponentially small magnitude unless
the method changes the representation of the finite-regulator integral.
\end{proof}

\section{Fermion Determinants}

For lattice fermions with quadratic action
\[
  \bar\psi D[U]\psi ,
\]
Berezin integration gives
\[
  \int d\bar\psi\,d\psi\,e^{-\bar\psi D[U]\psi}
  =
  \det D[U].
\]
If \(D\) obeys a positivity relation, such as
\[
  \gamma_5D\gamma_5=D^\dagger
\]
together with flavor pairing, the determinant can be nonnegative.  The
single-flavor statement and the paired-flavor statement are different.

\begin{remark}[Determinant reality from \(\gamma_5\)-Hermiticity]
\label{rem:gamma5-hermiticity-determinant-positivity}
Let \(D\) be a finite lattice Dirac matrix and let \(\Gamma\) be an
involution with \(\Gamma^\dagger=\Gamma\) and \(\Gamma^2=1\).  If
\begin{equation}
  \Gamma D\Gamma=D^\dagger ,
  \label{eq:finite-gamma5-hermiticity}
\end{equation}
then \(\det D\in\mathbb R\).  For two degenerate flavors with fermion
integral \((\det D)^2\), the finite weight is nonnegative because the real
determinant is squared.  For a conjugate pair of flavors with determinant
\begin{equation}
  \det D\,\det D^\dagger=|\det D|^2,
  \label{eq:paired-flavor-positive-determinant}
\end{equation}
the finite-regulator weight is nonnegative before possible zero-mode
subtleties.  Thus a positivity claim must state the actual flavor and
operator pairing together with the formal \(\gamma_5\)-Hermiticity relation.
Taking determinants in \eqref{eq:finite-gamma5-hermiticity} gives
\[
  \det D^\dagger
  =
  \det(\Gamma D\Gamma)
  =
  \det(\Gamma)^2\det D
  =
  \det D .
\]
Since \(\det D^\dagger=\overline{\det D}\), the determinant is real.  A real
number can have either sign, so \(\gamma_5\)-Hermiticity alone does not give
a positive single-flavor measure.  For an even number of exactly degenerate
copies, \((\det D)^{2k}\geq0\).  The conjugate-pair identity
\eqref{eq:paired-flavor-positive-determinant} is the elementary determinant
identity \(\det D^\dagger=\overline{\det D}\).  If \(\det D=0\), the
nonnegative weight vanishes and sampling remains mathematically defined, but
algorithmic ergodicity and condition-number issues become separate finite
regulator questions.
\end{remark}

At finite chemical potential or in chiral gauge theories this positivity may
fail.  The finite regulator then has no declared positive probability measure
whose expectations are the target correlators without reweighting or contour
deformation.

The deterministic check
\[
  \texttt{calculation-checks/monte\_carlo\_sign\_problem\_checks.py}
\]
verifies the finite-\(N\) autocorrelation identity, the phase-reweighting
variance bound, and the determinant-reality-versus-positivity distinction in
small finite examples.

\section{Continuum Interpretation}

Monte Carlo data become QFT data only after the regulator, source
normalization, operator renormalization, scaling trajectory, and error limits
are specified.  A numerical estimate at finite lattice spacing is a statement
about \(d\nu_\Lambda\).  A continuum claim requires convergence of a family of
finite-regulator expectations in the topology needed for OS reconstruction,
Wightman reconstruction, or the intended observable sector.

\begin{openproblem}[Sign-problem-preserving continuum construction]
\label{op:sign-problem-continuum-construction}
Classify QFTs with complex Euclidean weights for which a nonperturbative
continuum limit can be constructed by a positive enlarged-variable measure,
an exact contour deformation, or another finite-regulator method with
controlled variance and locality.
\end{openproblem}
