\chapter{Continuum Limits and Scaling Windows}
\label{ch:constructive-continuum-limits-scaling-windows}

\ConventionsEuclidean

A regulated model becomes a continuum QFT only after its correlation
functions converge as distributions and the limiting Schwinger functions
satisfy the hypotheses needed for reconstruction.  The continuum limit is
therefore a statement about a family of measures, a scaling trajectory in
the regulated action, and a coordinate system for local observables.  The
object produced is the limiting system of Schwinger distributions together
with its reconstruction data.

Throughout this chapter \(D\) is the Euclidean spacetime dimension,
\(\Lambda_a=a\mathbb Z^D\), and \(a>0\) is the lattice spacing.  The
infinite-volume limit is taken before \(a\downarrow0\), except where a
finite-volume scaling limit is explicitly stated.

\section{Regulated Families and Schwinger Distributions}

For each \(a\), let \(\mathcal E_a\) be a space of lattice fields and let
\[
  d\mu_a(\phi)=Z_a^{-1}\exp[-S_a(\phi)]D\phi
\]
be the regulated expectation functional.  The notation \(D\phi\) denotes the
finite-dimensional Lebesgue or Berezin product before infinite volume; after
an infinite-volume limit it denotes the resulting cylindrical expectation,
understood through its values on local observables.

A scaling trajectory is a map
\[
  a\longmapsto c(a)
\]
from lattice spacing to the local coefficients appearing in \(S_a\).  These
coefficients include all regulator-dependent local terms.  In this
monograph the word counterterm means such a local term in the regulated
action.  It is part of the action defining the path integral, not a second
action added afterward.

Let \(O^I_a(n)\) be local lattice observables.  A renormalized smeared
observable is
\begin{equation}
  O^I_a(f)
  =
  a^D\sum_{n\in\mathbb Z^D}
  f(an)\,Z^I{}_J(a)\,O^J_a(n),
  \qquad f\in C_c^\infty(\mathbb R^D),
  \label{eq:continuum-limit-renormalized-smeared-observable}
\end{equation}
where \(Z^I{}_J(a)\) may mix all observables with the same quantum numbers
and dimension bounded by the chosen truncation.  A continuum Schwinger
distribution is the multilinear distribution
\begin{equation}
  S^{I_1\ldots I_k}(f_1,\ldots,f_k)
  =
  \lim_{a\downarrow0}
  \left\langle
    O^{I_1}_a(f_1)\cdots O^{I_k}_a(f_k)
  \right\rangle_a ,
  \label{eq:continuum-limit-schwinger-distribution}
\end{equation}
when the limit exists for all test functions in the stated topology.

The topology matters.  A useful constructive statement gives uniform bounds
on the regulated distributions in a negative Sobolev or Schwartz seminorm,
then proves convergence on a dense test-function space.  Pointwise
convergence of unsmeared lattice correlators at separated lattice sites is a
weaker statement; it does not determine contact terms and does not by itself
give an operator-valued distribution after reconstruction.

\section{Free Lattice Scalar as a Model Calculation}

The massive free scalar illustrates the exact normalization of the
continuum limit.  Let
\begin{equation}
  S_a(\phi)
  =
  \frac12 a^D\sum_{x\in a\mathbb Z^D}
  \left[
    \sum_{\mu=1}^D
    \frac{(\phi_{x+a\hat\mu}-\phi_x)^2}{a^2}
    +m_a^2\phi_x^2
  \right],
  \label{eq:continuum-limit-free-lattice-action}
\end{equation}
with \(m_a\to m>0\).  The infinite-volume covariance is
\begin{equation}
  C_a(x-y)
  =
  \int_{B_a}\frac{d^Dp}{(2\pi)^D}\,
  \frac{e^{\ii p\cdot(x-y)}}
       {\hat p_a^2+m_a^2},
  \qquad
  \hat p_a^2=\frac{4}{a^2}\sum_{\mu=1}^D
  \sin^2\!\left(\frac{ap_\mu}{2}\right),
  \label{eq:continuum-limit-free-lattice-covariance}
\end{equation}
where \(B_a=(-\pi/a,\pi/a]^D\) is the Brillouin zone.  Define the smeared
field
\[
  \phi_a(f)=a^D\sum_{x\in a\mathbb Z^D} f(x)\phi_x .
\]

\begin{proposition}[Massive free lattice scalar continuum limit]
\label{prop:free-lattice-scalar-continuum-limit}
If \(m_a\to m>0\), then for every \(f,g\in C_c^\infty(\mathbb R^D)\),
\begin{equation}
  \lim_{a\downarrow0}
  \langle\phi_a(f)\phi_a(g)\rangle_a
  =
  \int_{\mathbb R^D}\frac{d^Dp}{(2\pi)^D}\,
  \frac{\widehat f(-p)\widehat g(p)}{p^2+m^2}.
  \label{eq:continuum-limit-free-scalar-limit}
\end{equation}
All higher Schwinger functions converge to the Gaussian Wick contractions of
this covariance.
\end{proposition}

\begin{proof}
For the covariance,
\[
  \langle\phi_a(f)\phi_a(g)\rangle_a
  =
  \int_{B_a}\frac{d^Dp}{(2\pi)^D}\,
  \frac{\widehat f_a(-p)\widehat g_a(p)}
       {\hat p_a^2+m_a^2},
\]
where
\[
  \widehat f_a(p)=a^D\sum_{x\in a\mathbb Z^D} f(x)e^{-\ii p\cdot x}.
\]
Poisson summation gives
\[
  \widehat f_a(p)
  =
  \sum_{\ell\in\mathbb Z^D}
  \widehat f(p+2\pi \ell/a).
\]
Since \(f\) is smooth and compactly supported, \(\widehat f\) is rapidly
decreasing; hence \(\widehat f_a(p)\to\widehat f(p)\) uniformly on every
fixed compact momentum set and has uniform rapid-decay bounds after
periodization.  Also
\[
  \hat p_a^2=p^2+O(a^2|p|^4)
\]
on compact sets, so the integrand converges pointwise to the continuum
integrand.  Split the Brillouin zone into \(|p|\le R\) and its complement.
On \(|p|\le R\), dominated convergence applies.  On the complement, rapid
decay of the periodized test functions and positivity of
\(\hat p_a^2+m_a^2\) give a bound whose \(R\to\infty\) limit is uniform for
small \(a\).  This proves the two-point convergence.  The finite-dimensional
Gaussian measures are determined by their covariance, so Wick's theorem
gives convergence of every smeared \(k\)-point function.
\end{proof}

This proof is the prototype of a constructive continuum statement:
regulated objects are smeared first, the limiting object is a distribution,
and the proof supplies a uniform estimate that lets the limit pass through
the integral.

\section{Tree-Level Symanzik Expansion and Cutoff Effects}
\label{sec:continuum-limit-symanzik-cutoff-effects}

The phrase lattice artifact is precise only after one has chosen continuum
coordinates for both sources and operators.  At fixed lattice spacing the
theory has hypercubic symmetry, not full \(O(D)\) invariance.  The continuum
limit may recover \(O(D)\) invariance because hypercubic but non-\(O(D)\)
local operators are irrelevant along the scaling trajectory, or because they
have been canceled by improvement terms.  This is the content, in its
tree-level form, of the Symanzik expansion.

\begin{definition}[Symanzik expansion data]
\label{def:symanzik-expansion-data}
Fix a regulated lattice family, a scaling trajectory \(c(a)\), and
renormalized observable coordinates \(O_a^I(f)\).  A Symanzik expansion to
order \(a^r\) for a class of smeared correlators is a finite list of
continuum local distributions \(S_{(n)}\), built from local operators
allowed by the lattice symmetries, such that
\begin{equation}
  \left\langle
    O_a^{I_1}(f_1)\cdots O_a^{I_k}(f_k)
  \right\rangle_a
  =
  \sum_{0\le n<r} a^n
  S_{(n)}^{I_1\ldots I_k}(f_1,\ldots,f_k)
  +
  R_{r,a}(f_1,\ldots,f_k),
  \label{eq:symanzik-expansion-data}
\end{equation}
with a stated seminorm \(q\) on the test functions and constants
\(C,a_0>0\) such that
\[
  |R_{r,a}(f_1,\ldots,f_k)|
  \le C a^r\,q(f_1)\cdots q(f_k),
  \qquad 0<a<a_0,
\]
or with another explicitly stated remainder estimate.  In an interacting
theory this is a theorem only after the uniform distributional bounds,
operator-mixing coordinates, and irrelevant operator estimates have been
proved.  The formal list of allowed local operators alone is not a
continuum-limit proof.
\end{definition}

\begin{proposition}[Free scalar tree-level cutoff expansion]
\label{prop:free-scalar-tree-level-symanzik-expansion}
For the free lattice scalar of
\eqref{eq:continuum-limit-free-lattice-action}, write
\[
  \hat p_{\mu,a}=\frac{2}{a}\sin\!\left(\frac{ap_\mu}{2}\right),
  \qquad
  \hat p_a^2=\sum_{\mu=1}^D\hat p_{\mu,a}^2 .
\]
Assume
\begin{equation}
  m_a^2=m^2+a^2\delta m_2+O(a^4)
  \label{eq:symanzik-free-mass-trajectory}
\end{equation}
with \(m>0\).  Then for every \(f,g\in C_c^\infty(\mathbb R^D)\),
\begin{equation}
  \langle\phi_a(f)\phi_a(g)\rangle_a
  =
  \int_{\mathbb R^D}\frac{\dd^Dp}{(2\pi)^D}\,
  \frac{\widehat f(-p)\widehat g(p)}{p^2+m^2}
  +
  a^2
  \int_{\mathbb R^D}\frac{\dd^Dp}{(2\pi)^D}\,
  \widehat f(-p)\widehat g(p)
  \frac{\frac1{12}\sum_{\mu}p_\mu^4-\delta m_2}
       {(p^2+m^2)^2}
  +O(a^4),
  \label{eq:free-scalar-tree-level-symanzik-propagator}
\end{equation}
where the \(O(a^4)\) remainder is bounded by a finite Schwartz seminorm of
\(f\) and \(g\).  The term \(\sum_\mu p_\mu^4\) is hypercubic invariant but
not \(O(D)\)-invariant for \(D>1\); it is the leading rotation-breaking
cutoff artifact of the nearest-neighbor action.
\end{proposition}

\begin{proof}
The Taylor expansion of the lattice momentum is
\begin{equation}
  \hat p_{\mu,a}^2
  =
  p_\mu^2-\frac{a^2}{12}p_\mu^4
  +\frac{a^4}{360}p_\mu^6+O(a^6p_\mu^8)
  \label{eq:lattice-momentum-symanzik-expansion}
\end{equation}
on every compact momentum set.  Hence the inverse kernel has the form
\[
  \hat p_a^2+m_a^2
  =
  p^2+m^2
  +
  a^2\left(
    \delta m_2-\frac1{12}\sum_\mu p_\mu^4
  \right)
  +
  O(a^4(1+|p|^6)).
\]
For \(A=p^2+m^2\) and
\[
  B=\delta m_2-\frac1{12}\sum_\mu p_\mu^4,
\]
the elementary identity
\[
  \frac1{A+a^2B+O(a^4)}
  =
  \frac1A-a^2\frac{B}{A^2}+O(a^4)
\]
gives the displayed coefficient.  The replacement of the Brillouin-zone
integral by the integral over \(\mathbb R^D\), and the uniform control of
the high-momentum tail, are the same Poisson-summation and rapid-decay
estimates used in the proof of
Proposition~\ref{prop:free-lattice-scalar-continuum-limit}.  The additional
polynomial factors in \(p\) are harmless because \(\widehat f\) and
\(\widehat g\) are Schwartz functions.  Thus the expansion holds as a
distributional expansion on smeared fields, with the remainder bounded by a
Schwartz seminorm.
\end{proof}

\begin{remark}[Tree-level scalar improvement]
\label{rem:tree-level-scalar-symanzik-improvement}
Define an improved free scalar quadratic kernel by
\begin{equation}
  K_a^{\rm imp}(p)
  =
  m_a^2+\hat p_a^2+\frac{a^2}{12}
  \sum_{\mu=1}^D\hat p_{\mu,a}^4 .
  \label{eq:tree-level-improved-scalar-kernel}
\end{equation}
If \(m_a^2=m^2+O(a^4)\), then
\begin{equation}
  K_a^{\rm imp}(p)
  =
  m^2+p^2+O(a^4(1+|p|^6))
  \label{eq:tree-level-improved-scalar-kernel-expansion}
\end{equation}
on compact momentum sets.  Consequently the \(O(a^2)\) correction in
\eqref{eq:free-scalar-tree-level-symanzik-propagator} is canceled for the
two-point function.
By \eqref{eq:lattice-momentum-symanzik-expansion},
\[
  \hat p_a^2
  =
  p^2-\frac{a^2}{12}\sum_\mu p_\mu^4+O(a^4|p|^6),
  \qquad
  \sum_\mu\hat p_{\mu,a}^4
  =
  \sum_\mu p_\mu^4+O(a^2|p|^6).
\]
The \(a^2\sum_\mu p_\mu^4/12\) improvement term cancels the
nearest-neighbor \(a^2\) artifact, and the mass assumption removes the
\(\delta m_2\) part.  The conclusion for smeared two-point functions follows
from the previous proposition.
\end{remark}

\begin{remark}[Status in interacting lattice theories]
\label{rem:symanzik-interacting-status}
For an interacting lattice theory, a Symanzik expansion is not obtained by
writing all local operators compatible with hypercubic symmetry.  One must
also prove that the regulated correlators admit the expansion after
renormalized operator mixing, and that the remainder is controlled in the
distribution topology used for reconstruction.  Gauge theories add the
further requirement that improvement terms and operator coordinates be
gauge-invariant or be formulated inside a gauge-fixed BV/lattice framework.
\end{remark}

\section{Correlation Length and Scaling Window}

The lattice correlation length is a property of a regulated state.  In a
massive translation-invariant theory, a connected two-point function at
large separation behaves like
\[
  \langle O(n)O(0)\rangle_{\rm conn}
  \sim e^{-|n|/\xi_{\rm lat}},
\]
where \(\xi_{\rm lat}\) is measured in lattice units.  A massive continuum
limit with physical mass \(M>0\) requires
\begin{equation}
  \xi_{\rm lat}(a)\to\infty,
  \qquad
  a\,\xi_{\rm lat}(a)\to M^{-1}.
  \label{eq:continuum-limit-massive-scaling-window}
\end{equation}
For a scale-invariant continuum limit, \(a\xi_{\rm lat}(a)\to\infty\).
The phrase scaling window means that the lattice spacing, observation
length, and correlation length are separated in the order
\[
  a\ll r\lesssim a\xi_{\rm lat}(a),
\]
with the last inequality replaced by \(r\ll a\xi_{\rm lat}(a)\) when one is
studying critical short-distance scaling inside a massive deformation.

For the free lattice scalar, the zero-spatial-momentum pole gives an exact
correlation length.  The pole in the Euclidean time direction is
\[
  \frac{4}{a^2}\sinh^2\!\left(\frac{a\kappa_a}{2}\right)=m_a^2,
  \qquad
  \kappa_a=\frac{2}{a}\operatorname{arcsinh}\!\left(\frac{am_a}{2}\right).
\]
Thus
\begin{equation}
  \xi_{\rm lat}(a)=\frac{1}{a\kappa_a}
  =
  \frac{1}{2\operatorname{arcsinh}(am_a/2)},
  \qquad
  a\xi_{\rm lat}(a)\to m^{-1}.
  \label{eq:continuum-limit-free-scalar-correlation-length}
\end{equation}
The expansion
\[
  \kappa_a=m_a-\frac{a^2m_a^3}{24}+O(a^4m_a^5)
\]
shows explicitly how the lattice mass approaches the physical pole mass.

\section{Finite-Volume Scaling and Step-Scaling Coordinates}
\label{sec:finite-volume-scaling-step-scaling-coordinates}

Finite volume is an infrared regulator and can be used as a renormalized
coordinate on the continuum theory through finite-volume observables.  To
separate this coordinate from the ultraviolet cutoff, keep three lengths
separate:
\[
  a,\qquad L=Na,\qquad \xi_{\rm phys}=a\,\xi_{\rm lat}.
\]
Here \(N\) is the number of lattice spacings across the box, \(L\) is the
physical size, and \(\xi_{\rm phys}\) is the physical correlation length.
A continuum finite-volume observable is obtained by taking
\[
  a\to0,\qquad L\ \text{fixed},\qquad
  \text{renormalized relevant coordinates fixed}.
\]
The infinite-volume limit \(L\to\infty\) is a subsequent infrared limit.
Interchanging these limits is an additional theorem, not part of the
definition of the finite-volume continuum object.

\begin{definition}[Finite-size scaling data]
\label{def:finite-size-scaling-data}
Fix a block scale \(b>1\), lattice tori
\((a\mathbb Z/Na\mathbb Z)^D\) with \(N=b^n\), and an RG chart near a
fixed point with one relevant coordinate \(t\), irrelevant coordinates
\(w=(w_\alpha)\), and local operators \(O^I\) of RG dimensions
\(\Delta_I\).  Finite-size scaling data on a compact set
\(\mathcal K\subset\mathbb R\) of scaling variables consists of:
\begin{enumerate}[label=(\roman*)]
  \item the integer \(n=\log_b N\);
  \item an endpoint coordinate
  \[
    u_N=t\,N^{y_t}+E_N(t,w),
    \qquad |E_N(t,w)|\le C N^{-\omega}
  \]
  whenever \(tN^{y_t}\in\mathcal K\);
  \item an irrelevant-coordinate estimate
  \[
    \|w_N\|\le C N^{-\omega}
  \]
  for some \(\omega>0\);
  \item endpoint kernels
  \(F^{I_1\ldots I_k}(u;z_1,\ldots,z_k)\), locally Lipschitz in \(u\),
  and renormalized separated-point kernels
  \(G_{N,t,w}^{I_1\ldots I_k}(z_1,\ldots,z_k)\), with
  \(z_i\in(\mathbb R/\mathbb Z)^D\) and \(z_i\ne z_j\), such that after
  \(n\) RG steps
  \[
    \left|
    N^{\Delta_{I_1}+\cdots+\Delta_{I_k}}
    G_{N,t,w}^{I_1\ldots I_k}(z_1,\ldots,z_k)
    -
    F^{I_1\ldots I_k}(u_N;z_1,\ldots,z_k)
    \right|
    \le C N^{-\omega}
  \]
  uniformly on compact subsets of the configuration space with pairwise
  separation bounded below.
\end{enumerate}
The constants \(C\) may depend on the compact set of noncoincident
configuration points and on the bounded microscopic set of irrelevant
coordinates, but not on \(N\).
\end{definition}

\begin{proposition}[Finite-size scaling from an RG chart]
\label{prop:finite-size-scaling-from-rg-chart}
Assume the finite-size scaling data of
Definition~\ref{def:finite-size-scaling-data}.  Then
\begin{equation}
  N^{\Delta_{I_1}+\cdots+\Delta_{I_k}}\,
  G_{N,t,w}^{I_1\ldots I_k}(z_1,\ldots,z_k)
  =
  F^{I_1\ldots I_k}
    (tN^{y_t};z_1,\ldots,z_k)
  +
  O(N^{-\omega})
  \label{eq:finite-size-scaling-rg-chart}
\end{equation}
uniformly for \(tN^{y_t}\in\mathcal K\).  If the continuum massive
coordinate is parameterized by a physical mass \(M\) through
\[
  t(a)=(aM)^{y_t}+O(a^{y_t+\epsilon})
\]
for some \(\epsilon>0\), then the scaling variable has the continuum
finite-volume limit
\begin{equation}
  t(a)N^{y_t}\to (ML)^{y_t}.
  \label{eq:finite-size-scaling-physical-variable}
\end{equation}
\end{proposition}

\begin{proof}
After \(n\) blocking steps, lengths have been divided by \(b^n=N\), and the
torus has order-one diameter in the blocked lattice units.  The
renormalized operator coordinate \(O^I\) contributes the factor
\(b^{-n\Delta_I}=N^{-\Delta_I}\).  The relevant coordinate at the endpoint
is \(u_N=tN^{y_t}+O(N^{-\omega})\), and the irrelevant coordinates have
norm \(O(N^{-\omega})\).  The endpoint estimate in the data gives the
correlator with \(F(u_N;z_1,\ldots,z_k)\) up to \(O(N^{-\omega})\).
Since the endpoint kernel is locally Lipschitz in \(u\) on
\(\mathcal K\), replacing \(u_N\) by \(tN^{y_t}\) changes the right-hand
side by another \(O(N^{-\omega})\).  This proves
\eqref{eq:finite-size-scaling-rg-chart}.  Finally, if \(N=L/a\) and
\(t(a)=(aM)^{y_t}+O(a^{y_t+\epsilon})\), then
\[
  t(a)N^{y_t}
  =
  (aM)^{y_t}(L/a)^{y_t}
  +
  O(a^\epsilon (L/a)^{y_t}a^{y_t})
  =
  (ML)^{y_t}+O(a^\epsilon L^{y_t}),
\]
which proves \eqref{eq:finite-size-scaling-physical-variable}.
\end{proof}

The proposition states exactly what must be proved before finite-size
scaling is a theorem: an RG chart, endpoint control, irrelevant decay, and
operator-coordinate stability.  The common formula
\(G_N=N^{-\sum\Delta_I}F(tN^{y_t})\) is the conclusion of these estimates,
not their replacement.

\begin{definition}[Step-scaling coordinate]
\label{def:step-scaling-coordinate}
Let \(\mathcal O_{a,L}\) be a finite-volume renormalized observable whose
regulator-level definition is gauge-invariant, or is a specified coordinate
in a gauge-fixed BV formulation, and whose continuum limit exists at fixed
physical size \(L\).  Suppose a real coordinate \(g_L\) is defined by an
injective normalization map
\[
  \mathcal O_{\rm cont}(L)=\Phi(g_L)
\]
on an interval of couplings.  For \(s>1\), the continuum step-scaling
function associated to this coordinate is
\begin{equation}
  \sigma_s(u)
  =
  \lim_{a/L\to0}
  g_{sL}(a)
  \quad
  \text{with }g_L(a)=u,
  \label{eq:step-scaling-coordinate-definition}
\end{equation}
provided the tuning \(g_L(a)=u\) and the continuum limit exist.  If the
cutoff effects are described by a Symanzik expansion, the numerical
extrapolation has the form
\begin{equation}
  g_{sL}(a)=\sigma_s(u)+
  \sum_{j=1}^{r-1} c_j(u,s)(a/L)^{\omega_j}
  +
  R_{r,a}(u,s),
  \label{eq:step-scaling-symanzik-extrapolation}
\end{equation}
with explicitly stated exponents and a stated bound on \(R_{r,a}\).
\end{definition}

This definition is the precise finite-volume version of a running coupling.
Different choices of \(\mathcal O\) define different schemes.  Scheme
dependence is not a defect: it is the coordinate dependence of the
renormalized theory.  What is scheme independent is the existence of the
continuum finite-volume observable, the consistency of the coordinate
changes where they overlap, and universal data such as fixed-point critical
exponents when the relevant theorem has been proved.

For lattice sizes not exactly of the form \(b^n\), the residual size
\(N/b^{\lfloor\log_bN\rfloor}\in[1,b)\) is an additional finite endpoint
coordinate.  It must be included in the scaling data or eliminated by an
explicit interpolation estimate; ignoring it is harmless only after that
bounded-coordinate dependence has been controlled.

\section{Operator Coordinates and Contact Terms}

Renormalized operator coordinates are part of the continuum-limit input.
For composite observables, the matrix \(Z^I{}_J(a)\) in
\eqref{eq:continuum-limit-renormalized-smeared-observable} may mix an
operator with derivatives of lower-dimensional operators and with the
identity.  The identity and derivative mixings are the lattice form of
contact terms.  They are visible only after smearing test functions and
studying coincident-point limits.

For example, the lattice square \(\phi_x^2\) in the free scalar model has
\[
  \langle\phi_x^2\rangle_a=C_a(0),
\]
which diverges for \(D\ge2\).  The Wick coordinate
\[
  :\!\phi_x^2\!:_a=\phi_x^2-C_a(0)
\]
has finite separated correlators and a well-defined smeared distributional
limit in the superrenormalizable dimensions where the corresponding
continuum Wick square exists.  If one changes the subtraction by a finite
local constant, the smeared operator changes by a multiple of
\(\int f(x)\dd^Dx\), namely by an identity contact coordinate.  Thus contact
terms are part of the coordinate chart on
local distributions.

\section{Osterwalder--Schrader Reconstruction Target}

\begin{definition}[Constructive continuum limit]
\label{def:constructive-continuum-limit}
A constructive continuum limit of a regulated family is a collection of
limiting Schwinger distributions satisfying Euclidean covariance,
permutation symmetry, reflection positivity, clustering, regularity, and the
growth assumptions required by the reconstruction theorem, together with the
OS-reconstructed Hilbert space, vacuum, fields, local domains, and
translation generators.
\end{definition}

Reflection positivity can pass to the limit by a closed-cone argument: if
\[
  \sum_{i,j}\overline{c_i}c_j
  S_{a}(\Theta F_i,F_j)\ge0
\]
for every finite set of positive-time test functions and the distributions
converge on these tests, then the limiting quadratic form is nonnegative.
Euclidean invariance is a separate limit theorem.  Hypercubic invariance at
fixed \(a\) becomes full Euclidean invariance only after one proves that
rotation-breaking local operators are irrelevant or have been tuned away.

The reconstruction target also includes growth bounds.  In Euclidean QFT the
Schwinger functions must be sufficiently controlled as tempered
distributions, and the analytic continuation to Wightman functions requires
the hypotheses discussed in
Chapter~\ref{ch:volume-iv-osterwalder-schrader-reconstruction}.  A continuum limit that
produces reflection-positive distributions but lacks the needed growth
control has not yet produced the full Lorentzian QFT object used elsewhere
in this monograph.

\section{Wilsonian Coordinates and Scaling Manifolds}

Let \(\mathcal B\) be a Banach space of local interactions near a fixed point
\(S_*\), and let \(\mathcal R_b:\mathcal U\subset\mathcal B\to\mathcal B\)
be a block-spin map with scale factor \(b>1\).  If \(\mathcal R_b(S_*)=S_*\)
and
\[
  D\mathcal R_b|_{S_*}\,v_i=b^{y_i}v_i ,
\]
then \(y_i>0\), \(y_i=0\), and \(y_i<0\) define relevant, marginal, and
irrelevant directions in this specified Banach norm and blocking scheme.
The classification is therefore tied to the chosen RG map and norm.

Suppose for simplicity that there is one relevant coordinate \(t\) with
eigenvalue \(b^{y_t}\), \(y_t>0\), and that all irrelevant coordinates are
controlled by a stable-manifold estimate.  Under \(N\) blocking steps,
\[
  t_N\approx b^{Ny_t}t_0.
\]
The correlation length in lattice units is the number of lattice spacings
needed until the running coordinate becomes order one, so
\[
  b^{Ny_t}|t_0|\sim1,
  \qquad
  \xi_{\rm lat}\sim b^N\sim |t_0|^{-1/y_t}.
\]
Thus the exponent relation
\begin{equation}
  \nu=\frac1{y_t}
  \label{eq:continuum-limit-nu-yt-relation}
\end{equation}
is a consequence of a controlled one-relevant-coordinate RG chart.  Without
the chart, the nonlinear estimates, and the identification of the
correlation length, the same formula is a scaling hypothesis rather than a
proof.

\section{Status of Scalar Examples}

Superrenormalizable scalar theories such as \(P(\phi)_2\) and
\(\phi^4_3\) have constructive continuum realizations with OS or Wightman
reconstruction under their stated hypotheses.  The existence theorem for
\(\phi^4_3\) as an interacting massive QFT is distinct from the assertion
that the three-dimensional Ising critical point is a conformal field theory
with the full expected operator spectrum and OPE.  The latter requires a
critical scaling limit with conformal covariance and operator-product
control, not only a massive constructive realization.

In four dimensions, perturbative renormalizability of \(\phi^4_4\) does not
settle nonperturbative ultraviolet existence.  The current rigorous status is
that no construction is known of a nonperturbative four-dimensional
UV-complete QFT whose perturbation series agrees with \(\phi^4_4\) to all
orders around the Gaussian fixed point.  Perturbative formal power series and
constructive continuum measures are different mathematical objects.

\section{Theorem Boundary}

The minimal data for a continuum-limit theorem are:
\[
\begin{gathered}
  \text{a regulated expectation functional for each }a,\\
  \text{a scaling trajectory }c(a),\\
  \text{renormalized observable coordinates }Z^I{}_J(a),\\
  \text{uniform distributional bounds and convergence},\\
  \text{reflection positivity, covariance, clustering, and growth control},\\
  \text{and an OS reconstruction of the limiting data.}
\end{gathered}
\]
For Wilsonian fixed-point arguments, the theorem must also specify the
function space, the blocking map, the fixed point, the relevant and
irrelevant projections, and estimates that control nonlinear remainders.
This is the line between a precise construction and a physically motivated
scaling picture.

\begin{openproblem}[Nonperturbative fixed points from Wilsonian maps]
\label{op:nonperturbative-wilsonian-fixed-points}
Develop a Wilsonian Banach-space framework for physically relevant
four-dimensional and three-dimensional QFTs in which fixed points,
renormalized operator coordinates, contact terms, scaling windows, and OS
reconstruction can all be controlled in the same theorem.  The framework
should make precise when a fixed-point argument is a proof and when it is a
well-supported construction principle.
\end{openproblem}
