\chapter{Lattice Fermions And Chiral Symmetry}
\label{ch:constructive-lattice-fermions-chiral-symmetry}

\ConventionsEuclidean

Fermions on a lattice require two separate pieces of structure.  The
variables integrated in the Euclidean functional integral are generators of a
finite Grassmann algebra, so the finite-cutoff integral is a Berezin
functional rather than a Borel measure.  The fields reconstructed after
Osterwalder--Schrader positivity are operator-valued distributions satisfying
graded locality.  The purpose of this chapter is to state the finite
algebraic regulator precisely, to identify where chiral symmetry is lost or
modified at nonzero lattice spacing, and to record the reflection-positive
fermion constructions that can serve as honest nonperturbative cutoffs.

\section{Finite Lattice Grassmann Integral}

Let \(\Lambda\subset a\mathbb Z^D\) be a finite lattice with a chosen ordering
of the pairs \((x,\alpha)\), where \(x\in\Lambda\) and
\(\alpha=1,\ldots,N_s\) is a spinor, flavor, and internal index.  The
fermionic field variables are odd generators
\[
  \psi_\alpha(x),\qquad \bar\psi_\alpha(x)
\]
of the complex Grassmann algebra
\(\bigwedge(\mathbb C^{2N_s|\Lambda|})\).  They are not points of a
configuration space.  A lattice fermion action quadratic in the fields is
specified by a finite matrix \(D\) with entries
\[
  D^{\alpha\beta}_{xy}:\mathbb C^{N_s|\Lambda|}
  \longrightarrow
  \mathbb C^{N_s|\Lambda|},
\]
and is written
\[
  S_D(\bar\psi,\psi)
  =
  \sum_{x,y\in\Lambda}
  \bar\psi_\alpha(x)D^{\alpha\beta}_{xy}\psi_\beta(y).
\]
The Berezin functional is normalized by
\[
  \int
  \prod_{(x,\alpha)}d\bar\psi_\alpha(x)d\psi_\alpha(x)\,
  \prod_{(x,\alpha)}\bar\psi_\alpha(x)\psi_\alpha(x)
  =
  1 .
\]
Changing this ordering changes both numerator and denominator by the same
sign in normalized correlation functions, but a fixed ordering must be
chosen before signs in reflection positivity or anomalies are meaningful.

\paragraph{Finite Berezin Gaussian identities.}
Let \(D\) be an invertible \(N\times N\) complex matrix and write
\[
  S_D=\sum_{i,j=1}^N\bar\psi_iD_{ij}\psi_j .
\]
With the Berezin ordering fixed above,
\[
  Z_D=
  \int\prod_i d\bar\psi_i\,d\psi_i\,e^{-S_D}
  =
  \det D .
\]
For every \(i,j\),
\[
  \frac{1}{Z_D}
  \int\prod_k d\bar\psi_k\,d\psi_k\,
  \bar\psi_j\psi_i\,e^{-S_D}
  =
  (D^{-1})_{ij}.
\]
More generally, for equal numbers of \(\psi\)'s and \(\bar\psi\)'s,
the normalized expectation is the determinant of the matrix of two-point
functions, with the sign fixed by the displayed order of the insertions.

Introduce independent odd sources \(\eta_i,\bar\eta_i\).  Completing the
square is an identity in the finite Grassmann algebra:
\[
  -\bar\psi D\psi+\bar\eta\psi+\bar\psi\eta
  =
  -(\bar\psi-\bar\eta D^{-1})D(\psi-D^{-1}\eta)
  +
  \bar\eta D^{-1}\eta .
\]
Berezin translation invariance gives
\[
  \int d\bar\psi\,d\psi\,
  e^{-\bar\psi D\psi+\bar\eta\psi+\bar\psi\eta}
  =
  \det D\,
  e^{\bar\eta D^{-1}\eta}.
\]
Differentiating with respect to \(\eta_i\) and \(\bar\eta_j\), using left
Grassmann derivatives in the fixed order, gives the two-point identity.
Higher identities follow by differentiating the same generating functional;
the exponential is quadratic in sources, so the coefficient of a monomial is
the corresponding determinant.

\section{Naive Dirac Operator And Doublers}

Let \(\gamma_\mu\), \(\mu=1,\ldots,D\), be Hermitian Euclidean gamma matrices,
\[
  \{\gamma_\mu,\gamma_\nu\}=2\delta_{\mu\nu}\mathbf 1 .
\]
On a periodic lattice, the naive massive Dirac operator is
\[
  D_{\rm naive}(p)
  =
  m\mathbf 1+\frac{\ii}{a}\sum_{\mu=1}^D\gamma_\mu\sin(ap_\mu),
  \qquad
  p_\mu\in(-\pi/a,\pi/a].
\]
For \(m\neq0\) it is invertible and
\begin{equation}
  D_{\rm naive}(p)^{-1}
  =
  \frac{
    m\mathbf 1-\frac{\ii}{a}\sum_\mu\gamma_\mu\sin(ap_\mu)}
  {m^2+a^{-2}\sum_\mu\sin^2(ap_\mu)} .
  \label{eq:lattice-naive-propagator-symbol}
\end{equation}
The massless symbol vanishes at every corner of the Brillouin zone:
\[
  p_\mu=\frac{\pi}{a}\epsilon_\mu,\qquad
  \epsilon_\mu\in\{0,1\}.
\]
The existence of these \(2^D\) zeros is the elementary doubling phenomenon.
The next proposition is a precise finite-dimensional version of the
topological obstruction in the translation-invariant nearest-continuum
setting.

\begin{proposition}[Brillouin-torus index cancellation]
\label{prop:lattice-fermion-brillouin-index-cancellation}
Let \(f=(f_1,\ldots,f_D):\mathbb T^D\to\mathbb R^D\) be a smooth periodic
map whose zeros are isolated and nondegenerate.  Consider the massless
translation-invariant Dirac symbol
\[
  D_f(p)=\ii\sum_{\mu=1}^D\gamma_\mu f_\mu(p).
\]
The chirality sign of a zero \(p_*\) is
\[
  \chi(p_*)=\operatorname{sign}\det
  \left(\frac{\partial f_\mu}{\partial p_\nu}(p_*)\right).
\]
Then
\[
  \sum_{f(p_*)=0}\chi(p_*)=0 .
\]
For \(f_\mu(p)=a^{-1}\sin(ap_\mu)\), the zero at the corner
\(\epsilon\in\{0,1\}^D\) has
\[
  \chi_\epsilon=(-1)^{\epsilon_1+\cdots+\epsilon_D},
\]
so the \(2^D\) naive species split equally between the two chirality signs.
\end{proposition}

\begin{proof}
For a regular value \(y\in\mathbb R^D\) of \(f\), define
\[
  I(y)=\sum_{f(p)=y}
  \operatorname{sign}\det
  \left(\frac{\partial f_\mu}{\partial p_\nu}(p)\right).
\]
The set of regular values is open and dense by Sard's theorem.  On a
connected component of regular values, \(I(y)\) is locally constant: if
\(y(s)\) is a small path of regular values and \(f(p_0)=y(0)\), the inverse
function theorem gives a unique smooth branch \(p(s)\) with
\(f(p(s))=y(s)\), and the sign of the nonzero Jacobian determinant cannot
change along the branch.  No signed zero can appear or disappear while the
value remains regular.

Since \(f(\mathbb T^D)\) is compact, choose \(y_\infty\) outside the image.
Then \(I(y_\infty)=0\).  Connect \(0\) to \(y_\infty\) by a piecewise smooth
path and perturb the path, keeping its endpoints fixed, so that it meets the
critical-value set only at isolated crossing parameters.  At such a crossing,
the standard local normal form for a birth--death critical point creates or
annihilates a pair of simple preimages with opposite Jacobian signs; hence
the signed count does not change.  Therefore \(I(0)=I(y_\infty)=0\).  This
is the degree-zero statement for a map from the torus to \(\mathbb R^D\).

For the naive symbol,
\[
  \frac{\partial f_\mu}{\partial p_\nu}
  =
  \delta_{\mu\nu}\cos(ap_\mu),
\]
and at \(p_\mu=\pi\epsilon_\mu/a\) the determinant is
\((-1)^{\epsilon_1+\cdots+\epsilon_D}\).
\end{proof}

\begin{remark}
The proposition proves the topological core of doubling for symbols of the
form \(D_f=\ii\gamma\cdot f\).  The full Nielsen--Ninomiya theorem allows a
broader class of local, translation-invariant, chirally symmetric lattice
Dirac operators.  Its proof is the same degree-theoretic obstruction after
one constructs the appropriate normalized vector field on the Brillouin
torus.  The hypotheses matter: dropping exact lattice chiral symmetry,
ultralocality, or the finite-component Dirac symbol changes the conclusion.
\end{remark}

\section{Wilson Fermions}

Wilson's construction changes the massless symbol by adding a spin-scalar
lattice Laplacian:
\begin{equation}
  D_W(p)
  =
  m\mathbf 1
  +\frac{\ii}{a}\sum_{\mu=1}^D\gamma_\mu\sin(ap_\mu)
  +\frac{r}{a}\sum_{\mu=1}^D(1-\cos(ap_\mu))\mathbf 1 ,
  \qquad r>0 .
  \label{eq:lattice-wilson-dirac-symbol}
\end{equation}
In position space, with gauge links \(U_\mu(x)\) in a unitary
representation of the gauge group,
\begin{equation}
\begin{aligned}
  (D_W\psi)(x)
  &=
  (m+rDa^{-1})\psi(x)                                      \\
  &\quad
  -\frac1{2a}\sum_\mu
  \Bigl[
    (r-\gamma_\mu)U_\mu(x)\psi(x+a\hat\mu)
    +(r+\gamma_\mu)U_\mu(x-a\hat\mu)^{-1}\psi(x-a\hat\mu)
  \Bigr].
\end{aligned}
  \label{eq:lattice-wilson-dirac-position}
\end{equation}
The Wilson term is gauge covariant and local at fixed \(a\).  It is not
chirally symmetric: when \(D\) is even and
\(\gamma_5=\ii^{D/2}\gamma_1\cdots\gamma_D\), the scalar Wilson term does
not anticommute with \(\gamma_5\).

\paragraph{Wilson lifting of the naive species.}
Let \(p=p_\epsilon+k\), where
\[
  (p_\epsilon)_\mu=\frac{\pi}{a}\epsilon_\mu,\qquad
  \epsilon_\mu\in\{0,1\},
\]
and \(ak_\mu\) is small.  Put
\[
  n_\epsilon=\epsilon_1+\cdots+\epsilon_D .
\]
Then the Wilson symbol has the expansion
\[
  D_W(p_\epsilon+k)
  =
  \left(m+\frac{2rn_\epsilon}{a}\right)\mathbf 1
  +\ii\sum_\mu(-1)^{\epsilon_\mu}\gamma_\mu k_\mu
  +O(a|k|^2).
\]
Thus the physical corner \(n_\epsilon=0\) has mass \(m\), while every
doubler corner has mass \(m+2rn_\epsilon/a\), which diverges in lattice
units as \(a\to0\) when \(m\) is held in the scaling window.

For \(p_\mu=\pi\epsilon_\mu/a+k_\mu\),
\[
  \sin(ap_\mu)=(-1)^{\epsilon_\mu}ak_\mu+O(a^3|k_\mu|^3),
\]
and
\[
  1-\cos(ap_\mu)
  =
  1-(-1)^{\epsilon_\mu}\cos(ak_\mu)
  =
  2\epsilon_\mu+O(a^2k_\mu^2).
\]
Substitution in \eqref{eq:lattice-wilson-dirac-symbol} gives the displayed
expansion.
The term \(2rn_\epsilon/a\) is present exactly at corners with
\(n_\epsilon>0\).  Thus the Wilson term distinguishes the physical corner
from the doubler corners by a regulator-scale mass shift while preserving the
linear Dirac behavior of the physical corner.

\section{Mid-link Reflection Positivity}

Reflection positivity for fermions is a statement about an antilinear
anti-automorphism of the Grassmann algebra and about the sign of the factor
coupling the two halves of the lattice.  It is not automatic from positivity
of a determinant.  We record the finite algebraic mechanism because this is
where many informal discussions lose the distinction between a Berezin
functional and a measure.

Let the reflection plane be the mid-link plane
\[
  x_0=\frac a2 .
\]
Let
\[
  \Lambda_+=\{x\in\Lambda:x_0\geq a\},
  \qquad
  \Lambda_-=\{x\in\Lambda:x_0\leq0\},
\]
and let \(\vartheta(t,\mathbf x)=(a-t,\mathbf x)\).  Choose the Euclidean
time gamma matrix so that \(\gamma_0^\dagger=\gamma_0\) and
\(\gamma_0^2=\mathbf 1\).  The reflection \(\Theta\) is defined on
positive-time generators by
\[
  \Theta\bigl(\psi_\alpha(x)\bigr)
  =
  (\gamma_0)_\beta{}^\alpha\bar\psi_\beta(\vartheta x),
  \qquad
  \Theta\bigl(\bar\psi_\alpha(x)\bigr)
  =
  (\gamma_0)_\alpha{}^\beta\psi_\beta(\vartheta x),
\]
extended antilinearly and with reversed order,
\[
  \Theta(AB)=\Theta(B)\Theta(A).
\]

Figure~\ref{fig:lattice-fermion-midlink-reflection} records the midlink
reflection used in the lattice-fermion reflection-positivity argument and the
induced exchange of fermion and antifermion variables under \(\Theta\).

\begin{figure}[H]
\centering
\begin{tikzpicture}[x=1.0cm,y=0.75cm]
  \draw[->] (-0.5,0)--(6.2,0) node[right] {\(x_0\)};
  \foreach \x/\lab in {0/\(0\),2/\(a\),4/\(2a\)} {
    \draw[fill=black] (\x,0) circle (1.5pt);
    \node[below] at (\x,-0.05) {\lab};
  }
  \draw[dashed,thick] (1,-1.0)--(1,1.0);
  \node[above] at (1,1.0) {\(x_0=a/2\)};
  \draw[thick,blue!60!black,-{Stealth[length=3pt]}] (2,0.35)--(0,0.35);
  \draw[thick,blue!60!black,-{Stealth[length=3pt]}] (0,-0.35)--(2,-0.35);
  \node[above] at (3,0.35) {\(\Lambda_+\)};
  \node[below] at (-0.05,-0.35) {\(\Lambda_-\)};
\end{tikzpicture}
\caption{The mid-link reflection plane lies between neighboring time slices.
The crossing hopping terms are the only terms that do not belong entirely to
one half of the lattice.}
\label{fig:lattice-fermion-midlink-reflection}
\end{figure}

\begin{definition}[Fermionic reflection-positive crossing kernel]
\label{def:lattice-fermion-rp-crossing-factor}
Let \(\mathcal A_+\) be the Grassmann algebra generated by positive-time
variables.  Choose an ordered family of homogeneous nilpotent positive-side
blocks
\[
  F_1,\ldots,F_M\in\mathcal A_+,
  \qquad
  |F_j|\in\mathbb Z/2,
\]
with disjoint Grassmann supports.  For each block fix the orientation phase
\(\tau_j\in U(1)\) determined by the chosen Berezin ordering, and write the
twisted reflected product as
\[
  \Theta(F_j)\boxtimes F_j
  =
  \tau_j\,\Theta(F_j)F_j .
\]
The phases are part of the reflection-positive datum: they are chosen so that
the saturation identity below has positive diagonal entries.  In the standard
pair ordering
\(\Theta(\chi_1)\chi_1\,\Theta(\chi_2)\chi_2\cdots\), one has
\(\tau_j=1\) for a singleton odd block and also for an ordered even block such
as \(\eta\bar\eta\).

A crossing kernel is reflection-positive if it can be written as
\[
  B_\times
  =
  \prod_{j=1}^M\bigl(1+c_j\,\Theta(F_j)\boxtimes F_j\bigr),
  \qquad c_j\geq0 .
\]
The value \(c_j=0\) is allowed; it records a positive-side block whose
reflected partner is supplied only when the observable itself contains that
block.
\end{definition}

\begin{lemma}[Finite Grassmann crossing-kernel criterion]
\label{prop:lattice-fermion-finite-rp-criterion}
Suppose a finite fermion functional has the form
\[
  \vev{A}
  =
  \frac1Z
  \int d\mu_-\,d\mu_+\,
  \Theta(e^{-S_+})\,e^{-S_+}\,B_\times\,A ,
\]
where \(S_+\in\mathcal A_+\), \(B_\times\) is a reflection-positive crossing
kernel in the sense of Definition~\ref{def:lattice-fermion-rp-crossing-factor},
and \(Z=\vev{1}>0\).  Then
\[
  \vev{\Theta(O)O}\geq0
\]
for every \(O\in\mathcal A_+\).
\end{lemma}

\begin{proof}
Set
\[
  \Psi=e^{-S_+}O .
\]
The bare Berezin functional is not an inner product on \(\mathcal A_+\).  With
one positive generator \(\chi\), and with \(\Theta(\chi)\) ordered before
\(\chi\),
\[
  \int d\Theta(\chi)\,d\chi\,1=0,
  \qquad
  \int d\Theta(\chi)\,d\chi\,\Theta(\chi)\chi=1 .
\]
The crossing kernel supplies the missing reflected-positive blocks.  The
orientation phases in Definition~\ref{def:lattice-fermion-rp-crossing-factor}
are fixed so that, for ordered block monomials \(F_I=\prod_{j\in I}F_j\),
\[
  \int d\mu_-\,d\mu_+\,
  \Theta(F_I)
  \left(\prod_{j\notin I}\Theta(F_j)\boxtimes F_j\right)
  F_K
  =
  \delta_{IK}.
\]
If a monomial cannot be completed to the full Berezin top form by the blocks
present in \(B_\times\), the same integral is zero; it is never a negative
diagonal contribution.  Expanding
\[
  \Psi=\sum_I z_I F_I
\]
in the ordered block basis and expanding \(B_\times\) gives
\[
  \int d\mu_-\,d\mu_+\,
  \Theta(\Psi)B_\times\Psi
  =
  \sum_I
  \left(\prod_{j\notin I}c_j\right)|z_I|^2 ,
\]
with the convention that a missing block with \(c_j=0\) gives weight zero.
All signs from reversing order under \(\Theta\) and from moving reflected
generators through positive generators are contained in the displayed twisted
products.  Since every coefficient is nonnegative, division by the positive
normalization \(Z\) preserves the inequality.
\end{proof}

\begin{proposition}[Wilson \(r=1\) mid-link reflection positivity]
\label{prop:lattice-fermion-midlink-free-rp}
For the gauge-covariant Wilson operator
\eqref{eq:lattice-wilson-dirac-position} with \(r=1\), unitary gauge links,
and reflection-invariant boundary conditions, the mid-link time reflection
\(x_0\mapsto a-x_0\) gives a reflection-positive finite Grassmann functional.
\end{proposition}

\begin{proof}
Let \(x_-=(0,\mathbf x)\), \(x_+=(a,\mathbf x)\), and
\[
  P_\pm=\frac12(1\pm\gamma_0).
\]
The cross-plane part of the Wilson action is exactly
\begin{equation}
  S_\times
  =
  -\frac1a\sum_{\mathbf x}
  \left[
    \bar\psi(x_-)P_-U_0(x_-)\psi(x_+)
    +
    \bar\psi(x_+)P_+U_0(x_-)^{-1}\psi(x_-)
  \right],
  \label{eq:wilson-midlink-crossing-action}
\end{equation}
because \(r-\gamma_0=2P_-\) and \(r+\gamma_0=2P_+\).  Reflection sends the
link \(U_0(x_-)\) to the inverse link on the reflected bond, and it exchanges
the two displayed bilinears.  Choose orthonormal bases of the spin-color
ranges of \(P_-\) and \(P_+\).  Since \(U_0(x_-)\) is unitary, the positive
linear forms obtained from
\[
  P_-U_0(x_-)\psi(x_+),
  \qquad
  \bar\psi(x_+)P_+U_0(x_-)^{-1}
\]
are again orthonormal odd generators after a unitary change of positive-side
Grassmann coordinates.  In that ordered basis
\[
  e^{-S_\times}
  =
  \prod_{\mathbf x,j}
  \left(1+\frac1a\,\Theta(F^-_{\mathbf x,j})\boxtimes F^-_{\mathbf x,j}\right)
  \left(1+\frac1a\,\Theta(F^+_{\mathbf x,j})\boxtimes F^+_{\mathbf x,j}\right).
\]
The coefficients are nonnegative and the projectors have no negative
eigenvalues.  All terms lying wholly on \(\Lambda_+\) are put in \(S_+\), and
their reflected partners are put in \(\Theta(S_+)\).  The finite
crossing-kernel criterion therefore gives
\[
  \vev{\Theta(O)O}\geq0
\]
for all \(O\in\mathcal A_+\).  The same calculation shows why the Wilson
coefficient matters: at \(r=0\) the time-crossing matrix is proportional to
\(\gamma_0\), whose positive and negative eigenvalues cannot be made into a
nonnegative crossing kernel by a unitary change of variables.
\end{proof}

The proposition is the reflection statement used in this chapter.  We do not
claim here that the mid-link cut is the only possible fermionic reflection
construction.  In particular, the common Fourier shortcut for a site
reflection that drops the \(\sin(ap_0)\) numerator as odd is not a proof at
nonzero reflected separation: the sine numerator can multiply the odd sine
part of the Fourier phase.  A site or diagonal reflection would require its
own complete finite Gram-matrix test, or a fully specified counterexample.

\section{Staggered Fermions And A Reflection-positive Thirring Coupling}

In two dimensions choose \(\gamma_0=\sigma_3\), \(\gamma_1=\sigma_1\).
The staggered one-component field \(\eta(x),\bar\eta(x)\) has kinetic matrix
\[
  M(x,x+a\hat\mu)=\frac{1}{2a}\eta_\mu(x),
  \qquad
  M(x,x-a\hat\mu)=-\frac{1}{2a}\eta_\mu(x-a\hat\mu),
\]
with
\[
  \eta_0(x)=1,\qquad \eta_1(x)=(-1)^{x_0/a}.
\]
The action is
\[
  S_{\rm stag}
  =
  \sum_{x,y}\bar\eta(x)M(x,y)\eta(y)
  +
  m\sum_x\bar\eta(x)\eta(x).
\]
The mid-link reflection acts by
\[
  \Theta(\eta(x))=(-1)^{x_1/a}\bar\eta(\vartheta x),
  \qquad
  \Theta(\bar\eta(x))=(-1)^{x_1/a}\eta(\vartheta x),
\]
with order reversal.

\begin{proposition}[Staggered Thirring reflection positivity]
\label{prop:lattice-staggered-thirring-rp}
Add the nearest-neighbor four-fermion interaction
\[
  \Delta S
  =
  U\sum_{x,\mu}
  \bar\eta(x)\eta(x+a\hat\mu)\bar\eta(x+a\hat\mu)\eta(x).
\]
For \(U\geq0\), the two-dimensional staggered fermion functional with action
\(S_{\rm stag}+\Delta S\) is reflection positive for the mid-link time
reflection.
\end{proposition}

\begin{proof}
The free staggered kinetic and mass terms have the mid-link crossing form
already described for the nearest-neighbor fermion functional: all terms
strictly on the positive half-lattice are placed in \(S_+\), their reflected
partners are placed in \(\Theta(S_+)\), and the hopping terms crossing the
reflection plane have a positive crossing-kernel expansion.

It remains to check the quartic interaction at the reflection plane.  Links
entirely inside one half again belong to \(S_+\) or to \(\Theta(S_+)\).  The
only interaction terms crossing the plane are, for \(x\) on the positive
boundary,
\[
  U\,\bar\eta(\vartheta x)\eta(x)\bar\eta(x)\eta(\vartheta x).
\]
Using the displayed reflection rule and the order reversal, this monomial is
the reflected-positive bilinear in the even positive-side generator
\(F_x=\eta(x)\bar\eta(x)\).  The orientation phase in
Definition~\ref{def:lattice-fermion-rp-crossing-factor} is fixed once for this
ordered even block.  With the displayed ordering it is \(+1\), so
\[
  e^{-\Delta S_{\times}}
  =
  \prod_{x\in P_+}
  \left(
    1+U\,\Theta(F_x)\boxtimes F_x
  \right),
\]
because the square of each Grassmann quartic monomial vanishes.  The
different boundary links involve distinct Grassmann generators, so the finite
product expands into the positive homogeneous-block cone with nonnegative
coefficients precisely when \(U\geq0\).  The full Boltzmann factor is
therefore in the form required by
Definition~\ref{def:lattice-fermion-rp-crossing-factor}, and the finite
crossing-kernel criterion,
Proposition~\ref{prop:lattice-fermion-finite-rp-criterion}, applies.
\end{proof}

\section{Ginsparg--Wilson Relation}

The Wilson term removes doublers by breaking the naive chiral symmetry.
Ginsparg--Wilson fermions replace the naive anticommutation relation by a
finite-cutoff identity.  We keep the normalization parameter visible because
the overlap operator below carries the domain-wall height \(\rho\).  Let
\(R>0\).  A lattice Dirac operator \(D\) satisfies the scalar
Ginsparg--Wilson relation if
\begin{equation}
  \gamma_5D+D\gamma_5=2R\,D\gamma_5D .
  \label{eq:lattice-ginsparg-wilson}
\end{equation}
Assume also \(\gamma_5\)-Hermiticity,
\[
  D^\dagger=\gamma_5D\gamma_5 .
\]
The common form \(\gamma_5D+D\gamma_5=aD\gamma_5D\) is the special case
\(R=a/2\); for the overlap convention used below, \(R=a/(2\rho)\).

\begin{lemma}[Exact lattice chiral symmetry and the finite Berezinian]
\label{lem:lattice-gw-chiral-jacobian}
Let \(D\) satisfy \eqref{eq:lattice-ginsparg-wilson} and
\(\gamma_5\)-Hermiticity.  The finite action
\[
  S=\bar\psi D\psi
\]
is invariant under
\[
  \delta\psi=\varepsilon\,\gamma_5(1-2RD)\psi,
  \qquad
  \delta\bar\psi=\varepsilon\,\bar\psi\gamma_5 .
\]
If the finite spinor space has \(\operatorname{Tr}\gamma_5=0\), the Berezin
measure transforms by
\[
  d\bar\psi'\,d\psi'
  =
  \exp\!\left(
    -2\varepsilon\,
    \operatorname{Tr}\gamma_5(1-RD)
    +O(\varepsilon^2)
  \right)
  d\bar\psi\,d\psi .
\]
The integer
\begin{equation}
  \operatorname{index}(D)
  =
  \operatorname{Tr}\gamma_5(1-RD)
  \label{eq:lattice-gw-index}
\end{equation}
equals the chirality difference \(n_+-n_-\) of the zero modes of \(D\).
\end{lemma}

\begin{proof}
The variation of the action is
\[
\begin{aligned}
  \delta S
  &=
  \varepsilon\,\bar\psi\gamma_5D\psi
  +
  \varepsilon\,\bar\psi D\gamma_5(1-2RD)\psi        \\
  &=
  \varepsilon\,\bar\psi
  \bigl(\gamma_5D+D\gamma_5-2R D\gamma_5D\bigr)\psi
  =
  0 .
\end{aligned}
\]
For odd variables, the Berezin measure transforms with the inverse
determinant.  Therefore
\[
  \log J
  =
  -\varepsilon\operatorname{Tr}\gamma_5(1-2RD)
  -\varepsilon\operatorname{Tr}\gamma_5
  +O(\varepsilon^2).
\]
When \(\operatorname{Tr}\gamma_5=0\), this is
\[
  2\varepsilon R\,\operatorname{Tr}\gamma_5D
  =
  -2\varepsilon\,\operatorname{Tr}\gamma_5(1-RD),
\]
which is the displayed Berezinian.

Define
\[
  \widehat\gamma_5=\gamma_5(1-2RD),
  \qquad
  \Gamma_5=\gamma_5(1-RD)=\frac12(\gamma_5+\widehat\gamma_5).
\]
\(\gamma_5\)-Hermiticity gives \(\widehat\gamma_5^\dagger=\widehat\gamma_5\).
Multiplying \eqref{eq:lattice-ginsparg-wilson} by \(\gamma_5\) on the left
gives
\[
  D+\gamma_5D\gamma_5=2R\gamma_5D\gamma_5D,
\]
and therefore \(\widehat\gamma_5^2=1\).  Hence
\(\operatorname{Tr}\Gamma_5=\frac12\operatorname{Tr}\widehat\gamma_5\) is an
integer.

It remains to identify the integer with zero-mode chirality.  The same
identity implies
\[
  D+D^\dagger=2R D^\dagger D,
\]
and its adjoint gives \(D+D^\dagger=2R DD^\dagger\); thus \(D\) is normal.
If \(D\phi=\lambda\phi\), then
\[
  D(\gamma_5\phi)=\overline{\lambda}\,\gamma_5\phi ,
\]
so nonreal eigenvalues are paired by \(\gamma_5\) and do not contribute to
\(\operatorname{Tr}\Gamma_5\).  The spectrum lies on the circle
\[
  \lambda+\overline{\lambda}=2R|\lambda|^2,
\]
whose real intersections are \(0\) and \(1/R\).  On \(\ker D\), \(\gamma_5\)
preserves the eigenspace and \(\Gamma_5=\gamma_5\), giving the contribution
\(n_+-n_-\).  On the \(1/R\) eigenspace, \(\Gamma_5=\gamma_5(1-R/R)=0\).
Thus \(\operatorname{Tr}\Gamma_5=n_+-n_-\).
\end{proof}

\section{Overlap Operator}

We now choose one convention and use it consistently.  Let
\(D_W^{(0)}\) be the massless Wilson operator, obtained from
\eqref{eq:lattice-wilson-dirac-symbol} by setting \(m=0\).  Define the
dimensionless Wilson kernel
\begin{equation}
  A_\rho=aD_W^{(0)}-\rho ,
  \qquad
  H_W(\rho)=\gamma_5A_\rho ,
  \label{eq:lattice-overlap-kernel-convention}
\end{equation}
where \(0<\rho\) is the domain-wall height.  The mass shift appears once:
either one works with \(D_W^{(0)}\) and subtracts \(\rho\), as here, or one
builds the negative mass into the Wilson operator and does not subtract it
again.  We use only the first convention.

Assume \(H_W(\rho)\) has no zero eigenvalues.  Let
\[
  \epsilon_\rho=\operatorname{sign}(H_W(\rho))
  =
  H_W(\rho)\bigl(H_W(\rho)^2\bigr)^{-1/2}.
\]
The massless overlap operator is
\begin{equation}
  D_{\rm ov}(\rho)
  =
  \frac{\rho}{a}
  \left(\mathbf 1+\gamma_5\epsilon_\rho\right).
  \label{eq:lattice-overlap-operator}
\end{equation}
The prefactor \(\rho/a\) fixes the physical branch to have unit Dirac slope.

\paragraph{Free branch window and normalization.}
For \(p=p_\epsilon+k\), \(p_{\epsilon,\mu}=\pi\epsilon_\mu/a\), the kernel is
\[
  A_\rho(p_\epsilon+k)
  =
  (2rn_\epsilon-\rho)\mathbf 1
  +
  \ii a\sum_\mu(-1)^{\epsilon_\mu}\gamma_\mu k_\mu
  +
  O(a^2|k|^2),
\]
where \(n_\epsilon=\epsilon_1+\cdots+\epsilon_D\).  At the physical corner
\(n_\epsilon=0\),
\[
  \frac{A_\rho(k)}{\sqrt{A_\rho(k)^\dagger A_\rho(k)}}
  =
  -1+\frac{\ii a}{\rho}\gamma_\mu k_\mu+O(a^2|k|^2),
\]
so
\[
  D_{\rm ov}(\rho;k)
  =
  \ii\gamma_\mu k_\mu+O(a|k|^2).
\]
For \(0<\rho<2r\), every doubler corner has \(2rn_\epsilon-\rho>0\), so the
corresponding overlap symbol has a regulator-scale value \(2\rho/a\) rather
than a second light branch.  At \(\rho=2rn\) a doubler shell crosses the kernel
zero.  If one first put a negative mass into \(D_W\) and then also formed
\(\gamma_5(D_W-\rho/a)\), the corner mass would be shifted twice and this
branch count would change.

\paragraph{Overlap operator satisfies Ginsparg--Wilson.}
Put
\[
  V_\rho=\gamma_5\epsilon_\rho .
\]
Since \(\epsilon_\rho^\dagger=\epsilon_\rho\) and
\(\epsilon_\rho^2=\mathbf 1\), \(V_\rho\) is unitary and
\(V_\rho^\dagger=\gamma_5V_\rho\gamma_5\).  Thus
\[
  D_{\rm ov}(\rho)=\frac{\rho}{a}(1+V_\rho),
  \qquad
  D_{\rm ov}(\rho)^\dagger
  =
  \gamma_5D_{\rm ov}(\rho)\gamma_5 .
\]
Unitarity gives
\[
  D_{\rm ov}+D_{\rm ov}^\dagger
  =
  \frac{a}{\rho}D_{\rm ov}^\dagger D_{\rm ov}.
\]
Multiplying by \(\gamma_5\) on the left gives
\[
  \gamma_5D_{\rm ov}+D_{\rm ov}\gamma_5
  =
  \frac{a}{\rho}D_{\rm ov}\gamma_5D_{\rm ov},
\]
which is \eqref{eq:lattice-ginsparg-wilson} with \(R=a/(2\rho)\).

\paragraph{Overlap index as spectral asymmetry.}
For the overlap operator,
\[
  \operatorname{index}(D_{\rm ov}(\rho))
  =
  -\frac12\operatorname{Tr}\epsilon_\rho .
\]
Indeed, using \eqref{eq:lattice-overlap-operator},
\[
  \gamma_5\left(1-\frac{a}{2\rho}D_{\rm ov}(\rho)\right)
  =
  \frac12\gamma_5-\frac12\epsilon_\rho .
\]
The finite spinor-lattice trace of \(\gamma_5\) vanishes, so the
Ginsparg--Wilson index \eqref{eq:lattice-gw-index} is the spectral asymmetry
of the Hermitian Wilson kernel.  By the spectral proof above this same trace
is the chirality difference of the overlap zero modes; the real modes at
\(2\rho/a\) do not contribute.

Finite invertibility, locality, and continuum topology are separate
statements.  Invertibility of \(H_W(\rho)\) is enough to define the finite
matrix \eqref{eq:lattice-overlap-operator}.  Exponential locality,
\[
  \|D_{\rm ov}(x,y)\|\leq C e^{-c|x-y|/a},
\]
requires a uniform exclusion of near-zero kernel modes, such as an
admissibility condition on gauge plaquettes or a mobility-gap estimate in the
gauge ensemble.  Identifying the finite trace with the continuum topological
charge requires a further scaling-limit theorem relating admissible lattice
gauge fields, kernel spectral flow, and the continuum Chern class.

\section{Domain-wall Fermions And The Overlap Limit}

Domain-wall fermions realize the same chiral mechanism by adding a finite
fifth lattice direction \(s=1,\ldots,L_s\).  In Shamir's normalization the
five-dimensional operator is built from \(A_\rho=aD_W^{(0)}-\rho\) and the
four-dimensional chiral projectors \(P_\pm=(1\pm\gamma_5)/2\):
\[
\begin{aligned}
  (\mathcal D_{\rm DW}(m_f)\Psi)_s
  &=
  (1+A_\rho)\Psi_s
  -P_-\Psi_{s+1}
  -P_+\Psi_{s-1}                                      \\
  &\quad
  +m_f\bigl(\delta_{s,1}P_+\Psi_{L_s}
            +\delta_{s,L_s}P_-\Psi_1\bigr),
\end{aligned}
\]
with open fifth-direction hopping except for the boundary mass \(m_f\).  The
physical boundary fields are
\[
  q=P_-\Psi_1+P_+\Psi_{L_s},
  \qquad
  \bar q=\bar\Psi_1P_+ + \bar\Psi_{L_s}P_- .
\]
The Pauli--Villars quotient removes the heavy bulk determinant:
\[
  \det D_{\rm eff}^{(L_s)}(m_f)
  =
  \frac{\det\mathcal D_{\rm DW}(m_f)}
       {\det\mathcal D_{\rm DW}(1)} .
\]

Block Gaussian elimination along the fifth direction gives a transfer matrix
description.  In the standard Shamir range, the Hermitian transfer kernel is
\[
  K_\rho=H_W(\rho)\,(2+A_\rho)^{-1},
\]
assuming the displayed inverse exists and the spectrum of \(K_\rho\) lies in
\((-1,1)\).  Then
\[
  T_\rho=(1-K_\rho)(1+K_\rho)^{-1},
  \qquad
  \varepsilon_{L_s}(K_\rho)
  =
  \frac{1-T_\rho^{L_s}}{1+T_\rho^{L_s}},
\]
and the massless boundary Schur complement is
\begin{equation}
  D_{\rm eff}^{(L_s)}(0)
  =
  \frac{\rho}{a}
  \left(1+\gamma_5\varepsilon_{L_s}(K_\rho)\right).
  \label{eq:domain-wall-finite-ls-effective-overlap}
\end{equation}
For an eigenvalue \(\kappa\in(-1,1)\) of \(K_\rho\),
\[
  \varepsilon_{L_s}(\kappa)
  =
  \tanh\bigl(L_s\,\operatorname{artanh}\kappa\bigr).
\]
If \(|\kappa|\geq\delta>0\) throughout the spectrum, then
\[
  \left\|
    \varepsilon_{L_s}(K_\rho)-\operatorname{sign}(K_\rho)
  \right\|
  \leq
  2\left(\frac{1-\delta}{1+\delta}\right)^{L_s}.
\]
Consequently
\[
  \left\|
    D_{\rm eff}^{(L_s)}(0)-D_{\rm ov}(\rho)
  \right\|
  \leq
  \frac{2\rho}{a}
  \left(\frac{1-\delta}{1+\delta}\right)^{L_s},
\]
provided \(K_\rho\) and \(H_W(\rho)\) have the same sign.  This is the
residual-mass mechanism in finite algebraic form: finite \(L_s\) gives an
exponentially accurate sign function, and the \(L_s\to\infty\) limit gives the
overlap operator with the same single Wilson-height shift used in
\eqref{eq:lattice-overlap-kernel-convention}.

\section{Continuum Requirements}

A lattice fermion construction used as a nonperturbative QFT regulator must
specify the following data.
\[
\begin{gathered}
  \text{finite Grassmann algebra and Berezin ordering},\\
  \text{Dirac operator, gauge covariance, spin structure, and boundary
  conditions},\\
  \text{reflection positivity or a replacement Hilbert-space construction},\\
  \text{chiral symmetry, anomaly, and index convention},\\
  \text{locality estimates uniform enough for a scaling limit},\\
  \text{operator regularization and renormalization separate from the
  cutoff action}.
\end{gathered}
\]
For chiral gauge theories, the fermion determinant or Pfaffian is a section
of a determinant line over the finite lattice gauge-field space.  A
nonperturbative definition requires a gauge-invariant, local, and
reflection-compatible choice of phase.  Anomaly cancellation is the local
cohomological condition that makes such a choice possible perturbatively; it
is not by itself a complete nonperturbative construction of the interacting
continuum theory.

\begin{openproblem}[Chiral gauge lattice]
\label{op:chiral-gauge-lattice}
Construct four-dimensional interacting chiral gauge theories on the lattice
with locality, reflection positivity or an equivalent reconstruction route,
exact gauge invariance, anomaly cancellation, and a continuum limit with the
desired local operator algebra.
\end{openproblem}
